### 2016-02-04 2016/PAVT/conclusion.tex
Add a draft of paper

\section{Заключение}
В работе рассмотрен параллельный алгоритм глобального поиска, разработанный в рамках информационно статистического подхода. Предложен подход к распараллеливанию данного алгоритма, который может быть эффективен как в задачах с трудоемким критерием оптимизации, так и в задачах с относительно простым критерием. Проведены эксперименты с реализацией предложенного параллельного алгоритма на суперкомпьютере «Лобачевский» с использованием сопроцессоров Intel Xeon Phi. Результаты вычислительных экспериментов подтверждают высокую эффективность распараллеливания для Xeon Phi во всех классах рассмотренных задач.

### 2016-02-04 2016/PAVT/dimension_reduction.tex
Add a draft of paper

\section{Обобщенная схема редукция размерности}
Одним из подходов к решению многомерных задач глобальной оптимизации является сведение их к одномерным и использование эффективных одномерных алгоритмов глобального поиска к редуцированной задаче. В предыдущем разделе была изложена идея редукции размерности с использованием кривых Пеано. Ниже излагается обобщенный способ редукции размерности, комбинирующий использование разверток и схему вложенной (рекурсивной) оптимизации.
\subsection{Рекурсивная схема редукции размерности}
Схема рекурсивной оптимизации основана на известном [10] соотношении
 ,				(9)
которое позволяет заменить решение многомерной задачи (1) решением семейства одно-мерных подзадач, рекурсивно связанных между собой.
Введем в рассмотрение множество функций 
 ,				(10)
 , .		(11)
Тогда, в соответствии с соотношением (9), решение исходной задачи  сводится к решению одномерной задачи
 .					(12)
Однако при этом каждое вычисление значения одномерной функции   в некоторой фиксированной точке предполагает решение одномерной задачи минимизации  ,и так далее до вычисления   согласно (10).
\subsection{Блочная рекурсивная схема редукции размерности}
Для изложенной выше рекурсивной схемы предложено обобщение (блочная рекурсивная схема), которое комбинирует использование разверток и рекурсивной схемы с целью эффективного распараллеливания вычислений.
Рассмотрим вектор y как вектор блочных переменных
 ,
где i-я блочная переменная ui представляет собой вектор размерности   из последовательно взятых компонент вектора y, т.е.  ,  ,…, , причем  .
С использованием новых переменных основное соотношение многошаговой схемы (9) может быть переписано в виде
	 ,				(13)
где подобласти  , являются проекциями исходной области поиска D на подпространства, соответствующие переменным  .
Формулы, определяющие способ решения задачи на основе соотношений (13) в целом совпадают с рекурсивной схемой (10)(12). Требуется лишь заменить исходные перемен-ные  , на блочные переменные  . 
При этом принципиальным отличием от исходной схемы является тот факт, что в блочной схеме вложенные подзадачи
	 , ,			(14)
являются многомерными, и для их решения может быть применен способ редукции раз-мерности на основе кривых Пеано.
Число векторов и количество компонент в каждом векторе являются параметрами блочной многошаговой схемы и могут быть использованы для формирования подзадач с нужными свойствами. Например, если  , т.е.  , то блочная схема идентична исходной; каждая из вложенных подзадач является одномерной. А если  , т.е.  , то решение задачи эквивалентно ее решению с использованием единствен-ной развертки, отображающей [0,1] в D; вложенные подзадачи отсутствуют.

### 2016-02-04 2016/PAVT/introduction.tex
Add a draft of paper

\section{Введение}
Рассматриваются задачи многоэкстремальной оптимизации и параллельные методы их решения. Важной особенностью указанных задач является тот факт, что глобальный экстремум есть интегральная характеристика задачи, таким образом, его отыскание связано с построением покрытия области поиска и вычислением значений оптимизируемой функции во всех точках этого покрытия. На сложность решения задач рассматриваемого класса решающее влияние оказывает размерность: вычислительные затраты растут экспоненциально при ее увеличении. Использование простейших способов решения (таких, как пере-бор на равномерной сетке) является неприемлемым. Требуется применение более экономных методов, которые порождают в области поиска существенно неравномерную сетку, более плотную в окрестности глобального минимума и разреженную вдали от него [13]. Данная статья продолжает развитие информационно-статистического подхода к построению параллельных алгоритмов глобальной оптимизации, предложенного в ННГУ им. Н.И. Лобачевского, который описан в монографиях [4, 5].
В рамках обсуждаемого подхода решение многомерных задач сводится к решению набора связанных подзадач меньшей размерности. Соответствующая редукция основана на использовании разверток единичного отрезка вещественной оси на гиперкуб. Роль таких разверток играют непрерывные однозначные отображения типа кривой Пеано, называемые также кривыми, заполняющими пространство. Еще одним используемым механизмом снижения размерности решаемой задачи является схема вложенной (рекурсивной) оптимизации. Численные методы, позволяющие эффективно использовать аппарат таких отображений, детально разработаны и обоснованы в [4, 5].
Алгоритмы, развиваемые в рамках информационно-статистического подхода, основаны на предположении липшицевости оптимизируемого критерия, что является типичными для других методов (см., например, [2, 3). Предположение такого рода выполняется для многих прикладных задач, поскольку относительные вариации функций, характеризую-щих моделируемую систему, обычно не превышают некоторый порог, определяемый ограниченной энергией изменений в системе.
Использование современных параллельных вычислительных систем расширяет сферу применения методов глобальной оптимизации и, в то же время, ставит задачу эффективно-го распараллеливания процесса поиска. Именно поэтому разработка эффективных параллельных методов для численного решения задач многоэкстремальной оптимизации и со-здание на их основе программных средств для современных вычислительных систем является актуальной задачей. Особый интерес представляет разработка схем распараллеливания, позволяющих эффективно использовать ускорители вычислений, такие, как сопроцессор Intel Xeon Phi.
В рамках проводимого исследования мы будем предполагать, что время проведения одного испытания (вычисления значения функции в области поиска) может значительно отличаться в различных решаемых задачах. Это определяет разработку разных подходов к распараллеливанию в задачах с «легкими» и «сложными» критериями. В статье приведено описание универсального подхода к распараллеливанию алгоритма глобального по-иска, который охватывает оба этих случая. Указанный подход реализован в разработанной в ННГУ им. Н.И. Лобачевского параллельной программной системе решения задач глобальной оптимизации.

### 2016-02-04 2016/PAVT/main.tex
Add a draft of paper

\documentclass[a4paper]{article}
\usepackage{fontspec}
\usepackage[russian]{babel}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[a4paper]{geometry}
\usepackage{indentfirst}
\usepackage{color}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\setmainfont{CMU Serif}
\setsansfont{CMU Sans Serif}

\usepackage[backend=biber]{biblatex}
\addbibresource{refs.bib}

\newcommand{\tbd}[1]{
    \textcolor{red}{#1}
}
\newcommand{\tworowcell}[2]{
    \begin{tabular}{@{}c@{}}#1 \\ #2\end{tabular}
}
\renewcommand{\setminus}{\mathbin{\backslash}}

\begin{document}

\title{Реализация параллельного алгоритма поиска глобального экстремума функции на Intel Xeon Phi\
\footnote{
Исследование выполнено при финансовой поддержке РФФИ в рамках научного проекта № 16-31-00244 мол\_а «Параллельные методы решения вычислительно трудоемких задач глобальной оптимизации на гибридных кластерных системах»}}
\author{К.А. Баркалов, И.Г. Лебедев, В.В. Соврасов, А.В. Сысоев}
\date{}
\maketitle

\begin{abstract}
Рассмотрен параллельный алгоритм решения задач многоэкстремальной оптимизации. Описывается реализация алгоритма на современных вычислительных системах с использованием сопроцессора Xeon Phi. Рассматриваются два подхода к распараллеливанию алгоритма, учитывающие информацию о трудоемкости вычисления значений оптимизируемой функции. Приводятся результаты вычислительных экспериментов, полученные на суперкомпьютере «Лобачевский». Демонстрируется, что реализация для Xeon Phi опережает версию для CPU. Результаты подтверждают ускорение алгоритма с использованием Xeon Phi по сравнению с алгоритмом, реализованным только на CPU.
\par
\textit{Ключевые слова:} глобальная оптимизация, многоэкстремальные функции, редукция размерности, параллельные алгоритмы, Intel Xeon Phi.
\end{abstract}

\input{introduction}
\input{parallel_alg}
\input{dimension_reduction}
\input{xeon_phi_implementation}
\input{numerical_experiments}
\input{conclusion}
\printbibliography
\end{document}

### 2016-02-04 2016/PAVT/numerical_experiments.tex
Add a draft of paper

\section{Результаты численных экспериментов}
Вычислительные эксперименты проводились на узле суперкомпьютера «Лобачевский», установленного в ННГУ им. Н.И. Лобачевского, на сегменте под управлением опе-рационной системы семейства Windows (HPC Server 2008). Узел располагает двумя процессорами Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM, и двумя сопроцессорами Intel Xeon Phi 5110P. 
В работе [6] описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных мини-мумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п. Тестовые задачи, порождаемые данным генератором, характеризуются малым временем вычисления значений целевой функции. Поэтому с целью имитации вычислительной трудоемкости, присущей прикладным задачам оптимизации, расчет критерия был усложнен дополнительными вычислениями, не меняющими вид функции и рас-положение ее минимумов.

Рис. 1. Решение двумерной задачи

Для примера на рис. 1 приведены линии уровня двумерной функции, порождаемой GKLS-генератором. Темные точки соответствуют 464 испытаниям, потребовавшимся для решения данной задачи с точностью 0.01 по координате; при этом число итераций составило 58 (на каждой итерации проводилось 8 испытаний). Линии уровня наглядно демонстрируют многоэкстремальность задачи, а расположение точек – неравномерное покрытие, сгущающееся только в районе глобального минимума.
Ниже приведены результаты численного сравнения трех последовательных алгоритмов – DIRECT [7], DIRECTl [8] и алгоритма глобального поиска (АГП) (результаты рабо-ты первых двух алгоритмов приводятся по работе [9]). Численное сравнение проводилось на классах функций Simple и Hard размерности 4 и 5 из [9]. Глобальный минимум y* считался найденным, если алгоритм генерировал точку испытания yk в -окрестности глобального минимума, т.е.  . При этом размер окрестности выбирался (в соответствии с [9]) как  , здесь N – размерность решаемой задачи, a и b – границы области поиска D, параметр =106 при N=4 и =107 при N=5. При использовании метода АГП для класса Simple выбирался параметр r=4.5, для класса Hard  r=5.6; параметр по-строения кривой Пеано был фиксированный m10. Максимально допустимое число итераций составляло Kmax = 1 000 000. 
В таблице 1 отражено среднее число итераций kav, которые выполнил метод при решении серии задач из данных классов. Символ «>» отражает ситуацию, когда не все зада-чи класса были решены каким-либо методом. Это означает, что алгоритм был остановлен по причине достижения максимально допустимого числа итераций Kmax. В этом случае значение Kmax=1 000 000 использовалось при вычислении среднего значения числа итераций kav, что соответствует нижней оценке этого среднего значения. Количество нерешенных задач указано в скобках.
Таблица 1. Среднее число итераций
N	Класс функции	DIRECT	DIRECTl	АГП
4	Simple	>47282 (4)	18983	11953
	Hard	>95708 (7)	68754	25263
5	Simple	>16057 (1)	16758	15920
	Hard	>217215 (16)	>269064 (4)	>148342 (4)

Как видно из таблицы 1, АГП превосходит методы DIRECT и DIRECTl на всех классах задач по среднему числу итераций. При этом в классе 5-Hard каждый из методов решил не все задачи: DIRECT не решил 16 задач, DIRECTl и АГП – по 4 задачи.
\subsection{Задачи с трудоемким критерием}
Вначале приведем результаты экспериментов, которые соответствуют задачам с большим временем проведения одного поискового испытания. Оценим ускорение параллельного алгоритм глобального поиска (ПАГП), реализованного на CPU, в зависимости от использованного количества ядер p. В таблицах 2 и 3 приведено ускорение по времени S(p) и по итерациям s(p). Ускорение вычислено относительно однопоточного запуска.

Таблица 2. Ускорение по времени S(p) на CPU
p	N4	N5
	Simple	Hard	Simple	Hard
2	2.45	2.20	1.15	1.32
4	4.66	3.90	2.82	2.59
8	7.13	7.35	3.47	5.34

Таблица 3. Ускорение по итерациям s(p) на CPU
p	N4	N5
	Simple	Hard	Simple	Hard
2	2.51	2.26	1.19	1.36
4	5.04	4.23	3.06	2.86
8	8.58	8.79	4.22	6.56

Результаты экспериментов показывают значительное ускорение ПАГП при использовании CPU. Лучшие результаты получены при использовании всех ядер процессора.
Далее изложим результаты экспериментов, проведенных с использованием Intel Xeon Phi. Вначале рассмотрим эксперименты на одном Xeon Phi в режиме Offload. Варьировалось количество потоков на сопроцессоре, все остальные параметры совпадают с предыдущими запусками. Приведено (таблицы 4 и 5) ускорение относительно восьмипоточного запуска на центральном процессоре.
Таблица 4. Ускорение по времени S(p) на Phi
p	N4	N5
	Simple	Hard	Simple	Hard
60	0.54	1.02	1.07	1.61
120	0.55	1.17	1.05	2.61
240	0.51	1.06	1.07	4.17
Таблица 5. Ускорение по итерациям s(p) на Phi
p	N4	N5
	Simple	Hard	Simple	Hard
60	8.13	7.32	9.87	6.55
120	16.33	15.82	15.15	17.31
240	33.07	27.79	38.80	59.31

Результаты экспериментов показывают, что только на классе Simple при N=4 реализация на Xeon Phi медленнее, чем CPU реализация, на классе Hard при N=4 и на классе Sim-ple при N=5 версии для Xeon Phi и CPU демонстрирует примерно равную эффективность, а на классе Hard при N=5 версия для Xeon Phi существенно превосходит CPU-реализацию. Наибольшее преимущество реализация для Xeon Phi показывает при 120 потоках для че-тырехмерных задач и при 240 потоках для пятимерных. При этом на всех классах наблюдается значительное ускорение по итерациям, а также практически линейная масштабируемость от числа потоков.
Далее рассмотрим эксперименты с запуском сопроцессора в режиме MPI. Число уровней вложенности разбиваемой задачи равно двум. Число процессов, запущенных на со-процессоре, – 30 (таблицы 6 и 7) и 60(таблицы 8 и 9), что соответствует ситуации, когда у корня дерева имеется соответственно 30 и 60 потомков. Процессы, работающие на сопроцессоре, использовали OpenMP для параллельного вычисления значений функции. Ускорение приведено относительно восьмипоточного запуска на CPU.
Таблица 6. Ускорение по времени S(p) на Phi, 30 MPI процессов
p	N4	N5
	Simple	Hard	Simple	Hard
4	3,18	5,14	5,15	14,65
8	3,53	1,49	3,35	10,49
16	1,09	2,77	4,61	14,36
Таблица 7. Ускорение по итерациям s(p) на Phi, 30 MPI процессов
p	N4	N5
	Simple	Hard	Simple	Hard
4	11,19	19,19	9,27	27,38
8	13,97	3,63	8,24	23,85
16	1,29	3,56	8,28	22,12
Таблица 8. Ускорение по времени S(p) на Phi, 60 MPI процессов
p	N4	N5
	Simple	Hard	Simple	Hard
4	2,76	3,20	4,19	16,56
8	3,05	7,73	4,08	14,87
16	0,89	5,21	2,72	13,05
Таблица 9. Ускорение по итерациям s(p) на Phi, 60 MPI процессов
p	N4	N5
	Simple	Hard	Simple	Hard
4	10,36	3,34	6,01	26,04
8	9,04	28,81	7,68	25,02
16	1,27	28,66	7,19	30,38

Результаты экспериментов показывают, что запуск в режиме MPI значительно превосходит Offload режим по времени работы, однако проигрывает по числу итераций. Лучшие результаты для простого класса задач наблюдаются при использовании 30 процессов на сопроцессоре, а для сложного класса, при 60 процессах. Таким образом, можно сделать вывод, что задачи с большим временем работы, для которых важно число итераций мето-да оптимизации, лучше использовать запуск в режиме Offload, а для более быстро решаемых задач, лучше подходит режим MPI. 
\subsection{Задачи с легко вычисляемым критерием}
Рассмотрим теперь решение задач, в которых поисковые испытания проводятся быстро. Вначале приведем результаты OpenMP версии. В таблице 10 приведено ускорение относительно однопоточного запуска.
Таблица 10. Ускорение по времени S(p) на CPU
p	N4	N5
	Simple	Hard	Simple	Hard
2	1,09	1,09	0,56	0,57
4	1,10	1,06	0,74	0,66
8	0,90	1,11	0,45	1,77
16	0,66	0,92	0,76	0,66
32	0,29	0,38	0,22	0,55

Эксперименты показывают незначительное ускорение, а во многих случаях – и замедление при вычислениях только на центральном процессоре.
Далее рассмотрим ускорение при использовании Xeon Phi в режиме MPI. Результаты приведены относительно однопоточного запуска на CPU. В таблице 11 приведено ускорение по времени S(p) для запуска 30 процессов на Xeon Phi, в таблице 12 для 60 процессов.
Таблица 11. Ускорение по времени S(p) на Phi, 30 MPI процессов
p	N4	N5
	Simple	Hard	Simple	Hard
1	0,35	2,28	0,37	1,92
2	0,94	1,95	0,39	1,23
4	1,23	0,43	0,54	2,63
8	0,85	0,56	0,40	1,53

Таблица 12. Ускорение по времени S(p) на Phi, 60 MPI процессов
p	N4	N5
	Simple	Hard	Simple	Hard
1	0,59	2,56	1,22	3,30
2	0,89	0,87	1,29	2,76
4	1,09	1,03	0,38	5,32
8	0,70	0,78	0,69	2,19

Результаты экспериментов показывают ускорение на Xeon Phi, превосходящее ускорение на центральном процессоре. Лучшие результаты получены при использовании 30 процессов на Xeon Phi только для простых четырехмерных задач, все остальные задачи лучше всего решаются при использовании 60 процессов.
Теперь перейдем к эксперименту с шестимерными задачами простого класса. При решении шестимерной задачи с использованием одного процесса (вложенные подзадачи отсутствуют) и сохранением точности построения развертки m=10 требуется использовать числа с плавающей запятой расширенной точности, что ведет к значительному увеличению времени работы алгоритма. При этом в случае использования хотя бы одного уровня вложенных подзадач, задачи на каждом уровне будут решаться без подключения расширенной точности, что обеспечивает дополнительное преимущество блочной многошаговой схемы.
Рассмотрим результаты эксперимента с использованием Xeon Phi в режиме MPI, таблица 13. Число уровней вложенности разбиваемой задачи равно двум. Приведены результаты экспериментов для разбиений 3:3 (три переменных на первом уровне, три – на вто-ром) и 4:2 (4 переменных на первом уровне, две – на втором). Число процессов, запущенных на сопроцессоре равно 30 и 60, что соответствует тому, что у корня дерева имеется соответственно 30 и 60 потомков. Процессы, работающие на Xeon Phi, использовали OpenMP для параллельного вычисления значений функции. Ускорение приведено относительно однопоточного запуска на CPU.
Таблица 13. Ускорение по времени S(p) на Phi
p	30 процессов	60 процессов
	3:3	4:2	3:3	4:2
2	7,30	9,87	17,68	18,21
4	5,99	8,18	6,60	13,57
8	3,14	12,74	5,92	25,58
16	3,54	7,61	4,87	11,33
32	3,41	6,85	3,73	6,50

Результаты экспериментов показывают значительное ускорение запуска на сопроцессоре по сравнению с запуском только на центральном процессоре. Лучшее ускорение наблюдается при шестидесяти MPI процессах на Xeon Phi по 8 потоков на каждый процесс, время работы в 25 раз меньше, чем решение задачи только на центральном процессоре.

### 2016-02-04 2016/PAVT/parallel_alg.tex
Add a draft of paper

\section{Параллельный алгоритм глобального поиска}
Рассмотрим задачу поиска глобального минимума N-мерной функции (y) в гиперинтервале   Будем предполагать, что функция удовлетворяет условию Липшица с априори неизвестной константой L.
	 ,					(1)
	 .			(2)
Существует ряд способов адаптации эффективных одномерных алгоритмов для решения многомерных задач, см., например, методы диагонального [6] или симплексного [7] разбиения области поиска. В данной работе мы будем использовать подход, основанный на идее редукции размерности с помощью кривой Пеано y(x), непрерывно и однозначно отображающей отрезок вещественной оси [0,1] на n-мерный куб 
 .
Вопросы численного построения отображений типа кривой Пеано и соответствующая теория подробно рассмотрены в [4]. Здесь же отметим, что численно построенная раз-вертка является приближением к теоретической кривой Пеано с точностью порядка , где m – параметр построения развертки. Использование подобного рода отображений позволяет свести многомерную задачу к одномерной задаче

Важным свойством является сохранение ограниченности относительных разностей функции: если функция (y) в области D удовлетворяла условию Липшица, то функция (y(x)) на интервале [0,1] будет удовлетворять равномерному условию Гельдера 
 ,  ,
где константа Гельдера H связана с константой Липшица L соотношением 
 ,  .
Поэтому, не ограничивая общности, можно рассматривать минимизацию одномерной функции  ,  , удовлетворяющей условию Гельдера.
Рассматриваемый алгоритм решения данной задачи предполагает построение последовательности точек xk, в которых вычисляются значения минимизируемой функции zk = f(xk). Процесс вычисления значения функции (включающий в себя построение образа yk=y(xk)) будем называть испытанием, а пару (xk,zk) – результатом испытания. Множество пар {(xk,zk)}, 1kn составляют поисковую информацию, накопленную методом после про-ведения n шагов. В нашем распоряжении имеется   вычислительных элементов и в рамках одной итерации метода мы будем проводить p испытаний одновременно. Обозначим   общее число испытаний, выполненных после n параллельных итераций.
На первой итерации метода испытание проводится в произвольной внутренней точке x1 интервала [0,1]. Пусть выполнено   итераций метода, в процессе которых были про-ведены испытания в  k = k(n) точках xi, 1ik. Тогда точки   поисковых испытаний следующей  -ой итерации определяются в соответствии с правилами:
Шаг 1. Перенумеровать точки множества  , которое включает в себя граничные точки интервала [0,1], а также точки предшествующих испытаний, ниж-ними индексами в порядке увеличения значений координаты, т.е.

Шаг 2. Полагая  , вычислить величины 
 ,  ,	(3)
где   является заданным параметром метода, а  .
Шаг 3. Для каждого интервала  , вычислить характеристику в соответствии с формулами
 ,  ,				(4)
 ,  .			(5)
Шаг 4. Характеристики  , упорядочить в порядке убывания 
	 				(6)
и выбрать   наибольших характеристик с номерами интервалов  .
Шаг 5. Провести новые испытания в точках  , вычисленных по формулам
 ,  
 ,  	(7)
Алгоритм прекращает работу, если выполняется условие   хотя бы для одного номера  ; здесь   есть заданная точность. В качестве оценки глобально-оптимального решения задачи  выбираются значения 
 ,  				(8)
Теоретическое обоснование данного способа организации параллельных вычислений изложено в [5].

### 2016-02-04 2016/PAVT/xeon_phi_implementation.tex
Add a draft of paper

\section{Реализация на Xeon Phi}
В2012 году компания Intel представила первый сопроцессор с архитектурой Intel MIC (Intel® Many Integrated Core Architecture). Архитектура MIC позволяет использовать боль-шое количество вычислительных ядер архитектуры x86 в одном процессоре. В результате для параллельного программирования могут быть использованы стандартные технологии, такие как OpenMP и MPI. 
Архитектура Intel Xeon Phi поддерживает несколько режимов использования сопроцессора, которые можно комбинировать для достижения максимальной производительно-сти в зависимости от характеристик решаемой задачи. 
В режиме Offload процессы MPI выполняются только на CPU, а на сопроцессоре про-исходит запуск отдельных функций, аналогично использованию графических ускорителей.
В режиме MPI базовая система и каждый сопроцессор Intel Xeon Phi рассматриваются как отдельные равноправные узлы, и процессы MPI могут выполняться на центральных процессорах и сопроцессорах Xeon Phi в произвольных сочетаниях.
\subsection{Режим Offload}
Вначале рассмотрим ситуацию, когда проведение одного испытания является трудоемкой операцией. В этом случае ускоритель Xeon Phi может быть использован в режиме Offload для параллельного проведения сразу многих испытаний на одной итерации метода. Пересылки данных от CPU к Xeon Phi будут минимальны – требуется лишь передать на сопроцессор координаты точек испытаний, и получить обратно значения функции в этих точках. Функции, определяющие обработку результатов испытаний в соответствии с алгоритмом и требующие работы с большим объемом накопленной поисковой информацией, могут быть эффективно реализованы на CPU. Общая схема организации вычислений с использованием Xeon Phi будет следующий.
На CPU выполняются шаги 1 – 4 параллельного алгоритма глобального поиска из п. 2. При этом на каждой итерации происходит накопление координат точек испытания в буфере, и этот буфер передается на сопроцессор. На Xeon Phi выполняется параллельное вы-числение значений функции в этих точках (шаг 5 алгоритма). Используется OpenMP распараллеливание цикла, на каждой итерации которого происходит вычисление значений функции. По завершению испытаний происходит передача вычисленных значений функ-ции на CPU. 
\subsection{Режим MPI}
В случае, если проведение одного поискового испытания является относительно про-стой операцией, параллельное проведение многих итераций в режиме Offload не дает большого ускорения (сказывается влияние накладных расходов на передачу данных). Од-нако здесь можно увеличить вычислительную нагрузку на Xeon Phi, если применить блочную схему редукции размерности из п. 3.2, а для решения возникающих подзадач использовать сопроцессор в режиме MPI.
Для организации параллельных вычислений будем использовать небольшое (2-3) число уровней вложенности в блочной схеме, при котором исходная задача большой размерности разбивается на 2-3 вложенные подзадачи меньшей размерности. Тогда, применяя в блочной рекурсивной схеме (13) для решения вложенных подзадач (14) параллельный алгоритм глобальной оптимизации, мы получим схему параллельных вычислений с широ-кой степенью вариативности (например, можно варьировать количество процессоров на различных уровнях оптимизации, т.е. при решении подзадач по различным переменным  ).
Общая схема организации вычислений с использованием нескольких узлов кластера и нескольких сопроцессоров состоит в следующем. Процессы параллельной программы об-разуют дерево, соответствующее уровням вложенных подзадач, при этом вложенные под-задачи

при i1,…,M–2 решаются только с использованием CPU. Непосредственно в данных под-задачах вычислений значений оптимизируемой функции не происходит: вычисление значения функции   это решение задачи минимизации следующего уровня. Каждая подзадача решается в отдельном процессе; обмен данными осуществляется лишь между процессами-предками и процессами-потомками.
Подзадача последнего (M–1)-го уровня

отличается от всех предыдущих подзадач – в ней происходит вычисление значений оптимизируемой функции, т.к.  . Подзадачи этого уровня решаются на сопроцессоре, и каждое ядро сопроцессора будет решать свою подзадачу (M–1)-го уровня в отдельном MPI-процессе.
Самый простой вариант использования данной схемы будет соответствовать двухкомпонентному вектору распараллеливания  . Здесь   будет соответствовать числу MPI-процессов на CPU, а  числу MPI-процессов на Xeon Phi; тем самым общее количество процессов будет определяться как  .
Отметим, что синхронный параллельный алгоритм глобальной оптимизации, описанный в п.3, в сочетании с блочной схемой редукции размерности обладает существенным недостатком, связанным с возможными простоями всех процессов, кроме корневого. Простои могут возникать в том случае, если часть потомков некоторого процесса закончили решение своих подзадач и отправили данные родителю раньше остальных. Для преодоления данной проблемы был применен предложенный в [12] асинхронный вариант параллельного алгоритма. Конкретные детали реализация блочной схемы в сочетании с асинхронным алгоритмом описаны в [12].

### 2016-02-08 2016/PAVT/introduction.tex
Finish introduction.tex

@@
 На сложность решения задач рассматриваемого класса решающее влияние оказывает размерность: вычислительные затраты растут экспоненциально при ее увеличении.
-Использование простейших способов решения (таких, как пере-бор на равномерной сетке) является неприемлемым.
-Требуется применение более экономных методов, которые порождают в области поиска существенно неравномерную сетку, более плотную в окрестности глобального минимума и разреженную вдали от него [13].
+Использование простейших способов решения (таких, как перебор на равномерной сетке) является неприемлемым.
+Требуется применение более экономных методов, которые порождают в области поиска существенно неравномерную сетку, более плотную в окрестности глобального минимума и разреженную вдали от него \cite{evtushenko}\cite{pinter}\cite{sergKvaDiag}.
 Данная статья продолжает развитие информационно-статистического подхода к построению параллельных алгоритмов глобальной оптимизации, предложенного в ННГУ им.
 Н.И.
-Лобачевского, который описан в монографиях [4, 5].
+Лобачевского, который описан в монографиях \cite{strSergOptBook}\cite{strGergrParOptBook}.
+\par
 В рамках обсуждаемого подхода решение многомерных задач сводится к решению набора связанных подзадач меньшей размерности.
@@
 Еще одним используемым механизмом снижения размерности решаемой задачи является схема вложенной (рекурсивной) оптимизации.
-Численные методы, позволяющие эффективно использовать аппарат таких отображений, детально разработаны и обоснованы в [4, 5].
-Алгоритмы, развиваемые в рамках информационно-статистического подхода, основаны на предположении липшицевости оптимизируемого критерия, что является типичными для других методов (см., например, [2, 3).
-Предположение такого рода выполняется для многих прикладных задач, поскольку относительные вариации функций, характеризую-щих моделируемую систему, обычно не превышают некоторый порог, определяемый ограниченной энергией изменений в системе.
+Численные методы, позволяющие эффективно использовать аппарат таких отображений, детально разработаны и обоснованы в \cite{strSergOptBook}\cite{strGergrParOptBook}.
+\par
+Алгоритмы, развиваемые в рамках информационно-статистического подхода, основаны на предположении липшицевости оптимизируемого критерия, что является типичными для других методов (см., например, \cite{pinter}\cite{sergKvaDiag}).
+Предположение такого рода выполняется для многих прикладных задач, поскольку относительные вариации функций, характеризующих моделируемую систему, обычно не превышают некоторый порог, определяемый ограниченной энергией изменений в системе.
+\par
 Использование современных параллельных вычислительных систем расширяет сферу применения методов глобальной оптимизации и, в то же время, ставит задачу эффективно-го распараллеливания процесса поиска.
@@
 Особый интерес представляет разработка схем распараллеливания, позволяющих эффективно использовать ускорители вычислений, такие, как сопроцессор Intel Xeon Phi.
+\par
 В рамках проводимого исследования мы будем предполагать, что время проведения одного испытания (вычисления значения функции в области поиска) может значительно отличаться в различных решаемых задачах.
 Это определяет разработку разных подходов к распараллеливанию в задачах с «легкими» и «сложными» критериями.
-В статье приведено описание универсального подхода к распараллеливанию алгоритма глобального по-иска, который охватывает оба этих случая.
+В статье приведено описание универсального подхода к распараллеливанию алгоритма глобального поиска, который охватывает оба этих случая.
 Указанный подход реализован в разработанной в ННГУ им.

### 2016-02-08 2016/PAVT/parallel_alg.tex
Finish introduction.tex

@@
 \section{Параллельный алгоритм глобального поиска}
-Рассмотрим задачу поиска глобального минимума N-мерной функции (y) в гиперинтервале   Будем предполагать, что функция удовлетворяет условию Липшица с априори неизвестной константой L.
-	 ,					(1)
-	 .			(2)
-Существует ряд способов адаптации эффективных одномерных алгоритмов для решения многомерных задач, см., например, методы диагонального [6] или симплексного [7] разбиения области поиска.
-В данной работе мы будем использовать подход, основанный на идее редукции размерности с помощью кривой Пеано y(x), непрерывно и однозначно отображающей отрезок вещественной оси [0,1] на n-мерный куб
- .
-Вопросы численного построения отображений типа кривой Пеано и соответствующая теория подробно рассмотрены в [4].
-Здесь же отметим, что численно построенная раз-вертка является приближением к теоретической кривой Пеано с точностью порядка , где m – параметр построения развертки.
+Рассмотрим задачу поиска глобального минимума \(N\)-мерной функции \(\phi(y)\) в гиперинтервале \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
+Будем предполагать, что функция удовлетворяет условию Липшица с априори неизвестной константой \(L\).
+\begin{equation}
+\label{task}
+\phi(y^*)=min\{\phi(y):y\in D\}
+\end{equation}
+\begin{equation}
+\label{lip}
+|\phi(y_1)-\phi(y_2)|\leqslant L||y_1-y_2||,y_1,y_2\in D,0<L<\infty
+\end{equation}
+\par
+Существует ряд способов адаптации эффективных одномерных алгоритмов для решения многомерных задач, см., например, методы диагонального \cite{sergKvaDiaPaper} или симплексного \cite{zilinsk} разбиения области поиска.
+В данной работе мы будем использовать подход, основанный на идее редукции размерности с помощью кривой Пеано \(y(x)\), непрерывно и однозначно отображающей отрезок вещественной оси \([0,1]\) на \(n\)-мерный куб
+\begin{equation}
+\label{cube}
+\{y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\}=\{y(x):0\leqslant x\leqslant 1\}
+\end{equation}
+\par
+Вопросы численного построения отображений типа кривой Пеано и соответствующая теория подробно рассмотрены в \cite{strSergOptBook}.
+Здесь же отметим, что численно построенная развертка является приближением к теоретической кривой Пеано с точностью порядка \(2^{-m}\), где \(m\) –-- параметр построения развертки.
 Использование подобного рода отображений позволяет свести многомерную задачу к одномерной задаче
+\begin{equation}
+\label{oneDimTask}
+\phi(y^*)=\phi(y(x^*))=min\{\phi(y(x)):x\in [0,1]\}
+\end{equation}
+\par
+Важным свойством является сохранение ограниченности относительных разностей функции: если функция \(\phi(y)\) в области \(D
+\) удовлетворяла условию Липшица, то функция \(\phi(y(x))\) на интервале \([0,1]\) будет удовлетворять равномерному условию Гельдера
+\begin{equation}
+\label{holder}
+|\phi(y(x_1))-\phi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
+\end{equation}
+где константа Гельдера \(H\) связана с константой Липшица \(L\) соотношением 
+\begin{equation}
+H=4Ld\sqrt{N},d=max\{b_i-a_i:1\leqslant i\leqslant N\}
+\end{equation}
+\par
+Поэтому, не ограничивая общности, можно рассматривать минимизацию одномерной функции  \(f(x)=\phi(y(x)), x\in[0,1]\), удовлетворяющей условию Гельдера.
+\par
+Рассматриваемый алгоритм решения данной задачи предполагает построение последовательности точек \(x_k\), в которых вычисляются значения минимизируемой функции \(z_k = f(x_k)\).
+Процесс вычисления значения функции (включающий в себя построение образа \(y_k=y(x_k))\) будем называть испытанием, а пару \((x_k,z_k)\) --- результатом испытания.
+Множество пар \(\{(x_k,z_k)\}, 1\leqslant k\leqslant n\) составляют поисковую информацию, накопленную методом после проведения \(n\) шагов.
+В нашем распоряжении имеется \(p\geqslant 1\) вычислительных элементов и в рамках одной итерации метода мы будем проводить \(p\) испытаний одновременно.
+Обозначим \(k(n)\) общее число испытаний, выполненных после \(n\) параллельных итераций.
+\par
+На первой итерации метода испытание проводится в произвольной внутренней точке \(x_1\) интервала \([0,1]\).
+Пусть выполнено \(n\leqslant 1\) итераций метода, в процессе которых были проведены испытания в \(k = k(n)\) точках \(x_i, 1\leqslant i\leqslant k\).
+Тогда точки \(x^{k+1},\dotsc,x^{k+p}\) поисковых испытаний следующей \((n+1)\)-ой итерации определяются в соответствии с правилами:
+\par
+Шаг 1.
+Перенумеровать точки множества \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), которое включает в себя граничные точки интервала \([0,1]\), а также точки предшествующих испытаний, нижними индексами в порядке увеличения значений координаты, т.е.
+\begin{equation}
+0=x_0<x_1<\dotsc<x_{k+1}=1
+\end{equation}
+\par
+Шаг 2. Полагая \(z_i=f(x_i),1\leqslant i\leqslant k\), вычислить величины 
+\begin{equation}
+\label{step2}
+\mu=max\dfrac{|z_i-z_{i-1}|}{\Delta_i}, M=
+\end{equation}
+где \(r\) является заданным параметром метода, а \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
+\par
+Шаг 3. Для каждого интервала \((x_{i-1},x_i),1\leqslant i\leqslant k+1\), вычислить характеристику в соответствии с формулами
+\begin{equation}
+\label{step3_1}
+\end{equation}
+\begin{equation}
+\label{step3_2}
+\end{equation}
+\par
+Шаг 4. Характеристики \(R(i),1\leqslant i\leqslant k+1\), упорядочить в порядке убывания 
+\begin{equation}
+\label{step4}
+\end{equation}
+и выбрать   наибольших характеристик с номерами интервалов \(t_j,1\leqslant j\leqslant p\).
+\par
+Шаг 5. Провести новые испытания в точках \(x_{k+j},1\leqslant j\leqslant p\), вычисленных по формулам
+\begin{equation}
+\label{step5}
+\end{equation}
+\par
+Алгоритм прекращает работу, если выполняется условие \(\Delta_{t_j}\leqslant \varepsilon\) хотя бы для одного номера \(t_j,1\leqslant j\leqslant p\); здесь \(\varepsilon>0\) есть заданная точность.
+В качестве оценки глобально-оптимального решения задачи  выбираются значения
+\begin{equation}
 
-Важным свойством является сохранение ограниченности относительных разностей функции: если функция (y) в области D удовлетворяла условию Липшица, то функция (y(x)) на интервале [0,1] будет удовлетворять равномерному условию Гельдера
- ,  ,
-где константа Гельдера H связана с константой Липшица L соотношением 
- ,  .
-Поэтому, не ограничивая общности, можно рассматривать минимизацию одномерной функции  ,  , удовлетворяющей условию Гельдера.
-Рассматриваемый алгоритм решения данной задачи предполагает построение последовательности точек xk, в которых вычисляются значения минимизируемой функции zk = f(xk).
-Процесс вычисления значения функции (включающий в себя построение образа yk=y(xk)) будем называть испытанием, а пару (xk,zk) – результатом испытания.
-Множество пар {(xk,zk)}, 1kn составляют поисковую информацию, накопленную методом после про-ведения n шагов.
-В нашем распоряжении имеется   вычислительных элементов и в рамках одной итерации метода мы будем проводить p испытаний одновременно.
-Обозначим   общее число испытаний, выполненных после n параллельных итераций.
-На первой итерации метода испытание проводится в произвольной внутренней точке x1 интервала [0,1].
-Пусть выполнено   итераций метода, в процессе которых были про-ведены испытания в  k = k(n) точках xi, 1ik.
-Тогда точки   поисковых испытаний следующей  -ой итерации определяются в соответствии с правилами:
-Шаг 1.
-Перенумеровать точки множества  , которое включает в себя граничные точки интервала [0,1], а также точки предшествующих испытаний, ниж-ними индексами в порядке увеличения значений координаты, т.е.
-
-Шаг 2. Полагая  , вычислить величины 
- ,  ,	(3)
-где   является заданным параметром метода, а  .
-Шаг 3. Для каждого интервала  , вычислить характеристику в соответствии с формулами
- ,  ,				(4)
- ,  .			(5)
-Шаг 4. Характеристики  , упорядочить в порядке убывания 
-	 				(6)
-и выбрать   наибольших характеристик с номерами интервалов  .
-Шаг 5. Провести новые испытания в точках  , вычисленных по формулам
- ,  
- ,  	(7)
-Алгоритм прекращает работу, если выполняется условие   хотя бы для одного номера  ; здесь   есть заданная точность. В качестве оценки глобально-оптимального решения задачи  выбираются значения 
- ,  				(8)
-Теоретическое обоснование данного способа организации параллельных вычислений изложено в [5].
+\end{equation}
+\par
+Теоретическое обоснование данного способа организации параллельных вычислений изложено в \cite{strGergrParOptBook}.

### 2016-02-11 2016/PAVT/parallel_alg.tex
Finish chapter parallel_alg.tex

@@
 \label{lip}
-|\phi(y_1)-\phi(y_2)|\leqslant L||y_1-y_2||,y_1,y_2\in D,0<L<\infty
+|\phi(y_1)-\phi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
 \end{equation}
@@
 В данной работе мы будем использовать подход, основанный на идее редукции размерности с помощью кривой Пеано \(y(x)\), непрерывно и однозначно отображающей отрезок вещественной оси \([0,1]\) на \(n\)-мерный куб
-\begin{equation}
+\begin{displaymath}
 \label{cube}
-\{y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\}=\{y(x):0\leqslant x\leqslant 1\}
-\end{equation}
+\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
+\end{displaymath}
 \par
@@
 Использование подобного рода отображений позволяет свести многомерную задачу к одномерной задаче
-\begin{equation}
+\begin{displaymath}
 \label{oneDimTask}
-\phi(y^*)=\phi(y(x^*))=min\{\phi(y(x)):x\in [0,1]\}
-\end{equation}
+\phi(y^*)=\phi(y(x^*))=\min\{\phi(y(x)):x\in [0,1]\}
+\end{displaymath}
 \par
@@
 \) удовлетворяла условию Липшица, то функция \(\phi(y(x))\) на интервале \([0,1]\) будет удовлетворять равномерному условию Гельдера
-\begin{equation}
+\begin{displaymath}
 \label{holder}
 |\phi(y(x_1))-\phi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
-\end{equation}
+\end{displaymath}
 где константа Гельдера \(H\) связана с константой Липшица \(L\) соотношением 
-\begin{equation}
-H=4Ld\sqrt{N},d=max\{b_i-a_i:1\leqslant i\leqslant N\}
-\end{equation}
+\begin{displaymath}
+H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}
+\end{displaymath}
 \par
@@
 На первой итерации метода испытание проводится в произвольной внутренней точке \(x_1\) интервала \([0,1]\).
-Пусть выполнено \(n\leqslant 1\) итераций метода, в процессе которых были проведены испытания в \(k = k(n)\) точках \(x_i, 1\leqslant i\leqslant k\).
+Пусть выполнено \(n\geqslant 1\) итераций метода, в процессе которых были проведены испытания в \(k = k(n)\) точках \(x_i, 1\leqslant i\leqslant k\).
 Тогда точки \(x^{k+1},\dotsc,x^{k+p}\) поисковых испытаний следующей \((n+1)\)-ой итерации определяются в соответствии с правилами:
@@
 Перенумеровать точки множества \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), которое включает в себя граничные точки интервала \([0,1]\), а также точки предшествующих испытаний, нижними индексами в порядке увеличения значений координаты, т.е.
-\begin{equation}
+\begin{displaymath}
 0=x_0<x_1<\dotsc<x_{k+1}=1
-\end{equation}
+\end{displaymath}
 \par
@@
 \label{step2}
-\mu=max\dfrac{|z_i-z_{i-1}|}{\Delta_i}, M=
+\mu=\max_{1\leqslant i\leqslant k}\dfrac{|z_i-z_{i-1}|}{\Delta_i},
+\begin{matrix}
+    M = 
+    \left\{
+    \begin{matrix}
+    r\mu,\mu>0 \\
+    1,\mu=0
+    \end{matrix} \right.
+    \end{matrix}
 \end{equation}
@@
 \label{step3_1}
+R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M}
 \end{equation}
@@
 \label{step3_2}
+R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1
 \end{equation}
@@
 \label{step4}
+R(t_1)\geqslant R(t_2)\geqslant \dots \geqslant R(t_{k})\geqslant R(t_{k+1})
 \end{equation}
-и выбрать   наибольших характеристик с номерами интервалов \(t_j,1\leqslant j\leqslant p\).
+и выбрать \(p\) наибольших характеристик с номерами интервалов \(t_j,1\leqslant j\leqslant p\).
 \par
 Шаг 5. Провести новые испытания в точках \(x_{k+j},1\leqslant j\leqslant p\), вычисленных по формулам
+\begin{displaymath}
+x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},t_j=1,t_j=k+1
+\end{displaymath}
 \begin{equation}
 \label{step5}
+x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-\sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{\mu}\right]^N,1<t_j<k+1
 \end{equation}
@@
 \begin{equation}
-
+f_k^*=\min_{1\leqslant i \leqslant k}f(x_i), x_k^*=\argmin_{1\leqslant i \leqslant k}f(x_i)
 \end{equation}

### 2016-02-13 2016/PAVT/dimension_reduction.tex
Finished dimension_reduction.tex and xeon_phi_implementation.tex

@@
 \subsection{Рекурсивная схема редукции размерности}
-Схема рекурсивной оптимизации основана на известном [10] соотношении
- ,				(9)
-которое позволяет заменить решение многомерной задачи (1) решением семейства одно-мерных подзадач, рекурсивно связанных между собой.
+Схема рекурсивной оптимизации основана на известном \cite{gorodGrishOptBook} соотношении
+\begin{equation}
+\label{nestedScheme}
+\min{\phi(y):y\in D}=\min_{a_1\leqslant y_1\leqslant b_1}\min_{a_2\leqslant y_2\leqslant b_2}\dots\min_{a_1\leqslant y_N\leqslant b_N}\phi(y)
+\end{equation}
+которое позволяет заменить решение многомерной задачи (\ref{task}) решением семейства одно-мерных подзадач, рекурсивно связанных между собой.
+\par
 Введем в рассмотрение множество функций 
- ,				(10)
- , .		(11)
-Тогда, в соответствии с соотношением (9), решение исходной задачи  сводится к решению одномерной задачи
- .					(12)
-Однако при этом каждое вычисление значения одномерной функции   в некоторой фиксированной точке предполагает решение одномерной задачи минимизации  ,и так далее до вычисления   согласно (10).
+\begin{equation}
+\label{phiN}
+\phi_N(y_1,\dots,y_N)=\phi(y_1,\dots,y_N)
+\end{equation}
+\begin{equation}
+\phi_i(y_1,\dots,y_i)=\min_{a_{i+1}\leqslant y_{i+1} \leqslant b_{i+1}}\phi_{i+1}(y_1,\dots,y_i,y_{i+1}),1\leqslant i\leqslant N-1
+\end{equation}
+\par
+Тогда, в соответствии с соотношением (\ref{nestedScheme}), решение исходной задачи  сводится к решению одномерной задачи
+\begin{equation}
+\label{phiFirst}
+\phi_1(y_1^*)=\min\{\phi_1(y_1),y_1\in [a,b]\}
+\end{equation}
+\par
+Однако при этом каждое вычисление значения одномерной функции \(\phi_1(y_1)\) в некоторой фиксированной точке предполагает решение одномерной задачи минимизации \(\phi_2(y_1,y_2^*)=\min\{\phi(y_1,y_2):y_2\in [a_2,b_2]\}\), и так далее до вычисления \(\phi_N\) согласно (\ref{phiN}).
 \subsection{Блочная рекурсивная схема редукции размерности}
 Для изложенной выше рекурсивной схемы предложено обобщение (блочная рекурсивная схема), которое комбинирует использование разверток и рекурсивной схемы с целью эффективного распараллеливания вычислений.
-Рассмотрим вектор y как вектор блочных переменных
- ,
-где i-я блочная переменная ui представляет собой вектор размерности   из последовательно взятых компонент вектора y, т.е.  ,  ,…, , причем  .
-С использованием новых переменных основное соотношение многошаговой схемы (9) может быть переписано в виде
-	 ,				(13)
-где подобласти  , являются проекциями исходной области поиска D на подпространства, соответствующие переменным  .
-Формулы, определяющие способ решения задачи на основе соотношений (13) в целом совпадают с рекурсивной схемой (10)(12). Требуется лишь заменить исходные перемен-ные  , на блочные переменные  . 
+\par
+Рассмотрим вектор \(y\) как вектор блочных переменных
+\begin{displaymath}
+y=(y_1,\dots,y_N)=(u_1,u_2,\dots,u_M)
+\end{displaymath}
+где \(i\)-я блочная переменная \(u_i\) представляет собой вектор размерности \(N\) из последовательно взятых компонент вектора \(y\), т.е.
+\(u_1=(y_1,y_2,\dots,y_{N_1}),u_2=(y_{N_1+1},y_{N_1+2},\dots,y_{N_1+N_2}),\dots,u_M=(y_{N-N_M+1},y_{N-N_M+2},\dots,y_N)\), причем \(N_1+N_2+\dots+N_M=N\).
+\par
+С использованием новых переменных основное соотношение многошаговой схемы (\ref{nestedScheme}) может быть переписано в виде
+\begin{equation}
+\label{blockNested}
+\min\phi(y)_{y\in D}=\min_{u_1\in D_1}\min_{u_2\in D_2}\dots\min_{u_M\in D_M}\phi(y)
+\end{equation}
+где подобласти \(D_i,1\leqslant i\leqslant M\), являются проекциями исходной области поиска \(D\) на подпространства, соответствующие переменным \(u_1,1\leqslant i\leqslant M\).
+\par
+Формулы, определяющие способ решения задачи на основе соотношений (\ref{blockNested}) в целом совпадают с рекурсивной схемой (\ref{phiN})-(\ref{phiFirst}).
+Требуется лишь заменить исходные переменные \(y_i,1\leqslant i\leqslant N\), на блочные переменные \(u_1,1\leqslant i\leqslant M\).
+\par
 При этом принципиальным отличием от исходной схемы является тот факт, что в блочной схеме вложенные подзадачи
-	 , ,			(14)
-являются многомерными, и для их решения может быть применен способ редукции раз-мерности на основе кривых Пеано.
+\begin{equation}
+\label{subTasks}
+\phi_i(u_1,\dots,u_i)=\min_{u_{i+1}\in D_{i+1}}\phi_{i+1}(u_1,\dots,u_i,u_{i+1}),1\leqslant i\leqslant M-1
+\end{equation}
+являются многомерными, и для их решения может быть применен способ редукции размерности на основе кривых Пеано.
+\par
 Число векторов и количество компонент в каждом векторе являются параметрами блочной многошаговой схемы и могут быть использованы для формирования подзадач с нужными свойствами.
-Например, если  , т.е.
-, то блочная схема идентична исходной; каждая из вложенных подзадач является одномерной.
-А если  , т.е.
-, то решение задачи эквивалентно ее решению с использованием единствен-ной развертки, отображающей [0,1] в D; вложенные подзадачи отсутствуют.
+Например, если \(M=N\), т.е.
+\(y_i=u_i,1\leqslant i\leqslant N\), то блочная схема идентична исходной; каждая из вложенных подзадач является одномерной.
+А если \(M=1\), т.е.
+\(u=u_1=y\), то решение задачи эквивалентно ее решению с использованием единственной развертки, отображающей \([0,1]\) в \(D\); вложенные подзадачи отсутствуют.

### 2016-02-13 2016/PAVT/numerical_experiments.tex
Finished dimension_reduction.tex and xeon_phi_implementation.tex

@@
 Н.И.
-Лобачевского, на сегменте под управлением опе-рационной системы семейства Windows (HPC Server 2008).
+Лобачевского, на сегменте под управлением операционной системы семейства Windows (HPC Server 2008).
 Узел располагает двумя процессорами Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM, и двумя сопроцессорами Intel Xeon Phi 5110P.
-В работе [6] описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных мини-мумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.
+\par
+В работе \cite{gklsPaper} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.
 Тестовые задачи, порождаемые данным генератором, характеризуются малым временем вычисления значений целевой функции.
-Поэтому с целью имитации вычислительной трудоемкости, присущей прикладным задачам оптимизации, расчет критерия был усложнен дополнительными вычислениями, не меняющими вид функции и рас-положение ее минимумов.
-
-Рис. 1. Решение двумерной задачи
-
+Поэтому с целью имитации вычислительной трудоемкости, присущей прикладным задачам оптимизации, расчет критерия был усложнен дополнительными вычислениями, не меняющими вид функции и расположение ее минимумов.
+\begin{figure}[h!]
+    \centering
+		\includegraphics[width=0.45\textwidth]{isolines.png}
+		\caption{Решение двумерной задачи}\label{fig:isolines}
+\end{figure}
+\par
 Для примера на рис.
-1 приведены линии уровня двумерной функции, порождаемой GKLS-генератором.
+\ref{fig:isolines} приведены линии уровня двумерной функции, порождаемой GKLS-генератором.
 Темные точки соответствуют 464 испытаниям, потребовавшимся для решения данной задачи с точностью 0.01 по координате; при этом число итераций составило 58 (на каждой итерации проводилось 8 испытаний).
-Линии уровня наглядно демонстрируют многоэкстремальность задачи, а расположение точек – неравномерное покрытие, сгущающееся только в районе глобального минимума.
+Линии уровня наглядно демонстрируют многоэкстремальность задачи, а расположение точек --- неравномерное покрытие, сгущающееся только в районе глобального минимума.
+\par
 Ниже приведены результаты численного сравнения трех последовательных алгоритмов – DIRECT [7], DIRECTl [8] и алгоритма глобального поиска (АГП) (результаты рабо-ты первых двух алгоритмов приводятся по работе [9]).
 Численное сравнение проводилось на классах функций Simple и Hard размерности 4 и 5 из [9].
-Глобальный минимум y* считался найденным, если алгоритм генерировал точку испытания yk в -окрестности глобального минимума, т.е.
+Глобальный минимум \(y^*\) считался найденным, если алгоритм генерировал точку испытания yk в -окрестности глобального минимума, т.е.
 .
@@
 В таблице 1 отражено среднее число итераций kav, которые выполнил метод при решении серии задач из данных классов.
-Символ «>» отражает ситуацию, когда не все зада-чи класса были решены каким-либо методом.
+Символ «>» отражает ситуацию, когда не все задачи класса были решены каким-либо методом.
 Это означает, что алгоритм был остановлен по причине достижения максимально допустимого числа итераций Kmax.

### 2016-02-13 2016/PAVT/xeon_phi_implementation.tex
Finished dimension_reduction.tex and xeon_phi_implementation.tex

@@
 \section{Реализация на Xeon Phi}
-В2012 году компания Intel представила первый сопроцессор с архитектурой Intel MIC (Intel® Many Integrated Core Architecture).
-Архитектура MIC позволяет использовать боль-шое количество вычислительных ядер архитектуры x86 в одном процессоре.
+В 2012 году компания Intel представила первый сопроцессор с архитектурой Intel MIC (Intel® Many Integrated Core Architecture).
+Архитектура MIC позволяет использовать большое количество вычислительных ядер архитектуры x86 в одном процессоре.
 В результате для параллельного программирования могут быть использованы стандартные технологии, такие как OpenMP и MPI.
-Архитектура Intel Xeon Phi поддерживает несколько режимов использования сопроцессора, которые можно комбинировать для достижения максимальной производительно-сти в зависимости от характеристик решаемой задачи.
-В режиме Offload процессы MPI выполняются только на CPU, а на сопроцессоре про-исходит запуск отдельных функций, аналогично использованию графических ускорителей.
+\par
+Архитектура Intel Xeon Phi поддерживает несколько режимов использования сопроцессора, которые можно комбинировать для достижения максимальной производительности в зависимости от характеристик решаемой задачи.
+\par
+В режиме Offload процессы MPI выполняются только на CPU, а на сопроцессоре происходит запуск отдельных функций, аналогично использованию графических ускорителей.
+\par
 В режиме MPI базовая система и каждый сопроцессор Intel Xeon Phi рассматриваются как отдельные равноправные узлы, и процессы MPI могут выполняться на центральных процессорах и сопроцессорах Xeon Phi в произвольных сочетаниях.
@@
 В этом случае ускоритель Xeon Phi может быть использован в режиме Offload для параллельного проведения сразу многих испытаний на одной итерации метода.
-Пересылки данных от CPU к Xeon Phi будут минимальны – требуется лишь передать на сопроцессор координаты точек испытаний, и получить обратно значения функции в этих точках.
+Пересылки данных от CPU к Xeon Phi будут минимальны --- требуется лишь передать на сопроцессор координаты точек испытаний, и получить обратно значения функции в этих точках.
 Функции, определяющие обработку результатов испытаний в соответствии с алгоритмом и требующие работы с большим объемом накопленной поисковой информацией, могут быть эффективно реализованы на CPU.
 Общая схема организации вычислений с использованием Xeon Phi будет следующий.
+\par
 На CPU выполняются шаги 1 – 4 параллельного алгоритма глобального поиска из п.
@@
 При этом на каждой итерации происходит накопление координат точек испытания в буфере, и этот буфер передается на сопроцессор.
-На Xeon Phi выполняется параллельное вы-числение значений функции в этих точках (шаг 5 алгоритма).
+На Xeon Phi выполняется параллельное вычисление значений функции в этих точках (шаг 5 алгоритма).
 Используется OpenMP распараллеливание цикла, на каждой итерации которого происходит вычисление значений функции.
-По завершению испытаний происходит передача вычисленных значений функ-ции на CPU.
+По завершению испытаний происходит передача вычисленных значений функции на CPU.
 \subsection{Режим MPI}
-В случае, если проведение одного поискового испытания является относительно про-стой операцией, параллельное проведение многих итераций в режиме Offload не дает большого ускорения (сказывается влияние накладных расходов на передачу данных).
-Од-нако здесь можно увеличить вычислительную нагрузку на Xeon Phi, если применить блочную схему редукции размерности из п.
+В случае, если проведение одного поискового испытания является относительно простой операцией, параллельное проведение многих итераций в режиме Offload не дает большого ускорения (сказывается влияние накладных расходов на передачу данных).
+Однако здесь можно увеличить вычислительную нагрузку на Xeon Phi, если применить блочную схему редукции размерности из п.
 3.2, а для решения возникающих подзадач использовать сопроцессор в режиме MPI.
+\par
 Для организации параллельных вычислений будем использовать небольшое (2-3) число уровней вложенности в блочной схеме, при котором исходная задача большой размерности разбивается на 2-3 вложенные подзадачи меньшей размерности.
-Тогда, применяя в блочной рекурсивной схеме (13) для решения вложенных подзадач (14) параллельный алгоритм глобальной оптимизации, мы получим схему параллельных вычислений с широ-кой степенью вариативности (например, можно варьировать количество процессоров на различных уровнях оптимизации, т.е.
-при решении подзадач по различным переменным  ).
+Тогда, применяя в блочной рекурсивной схеме (\ref{blockNested}) для решения вложенных подзадач (\ref{subTasks}) параллельный алгоритм глобальной оптимизации, мы получим схему параллельных вычислений с широкой степенью вариативности (например, можно варьировать количество процессоров на различных уровнях оптимизации, т.е.
+при решении подзадач по различным переменным \(u_i\)).
+\par
 Общая схема организации вычислений с использованием нескольких узлов кластера и нескольких сопроцессоров состоит в следующем.
-Процессы параллельной программы об-разуют дерево, соответствующее уровням вложенных подзадач, при этом вложенные под-задачи
-
-при i1,…,M–2 решаются только с использованием CPU.
-Непосредственно в данных под-задачах вычислений значений оптимизируемой функции не происходит: вычисление значения функции   это решение задачи минимизации следующего уровня.
+Процессы параллельной программы образуют дерево, соответствующее уровням вложенных подзадач, при этом вложенные подзадачи
+\begin{displaymath}
+\phi_i(u_1,\dots,u_i)=\min_{u_{i+1}\in D_{i+1}}\phi_{i+1}(u_1,\dots,u_i,u_{i+1})
+\end{displaymath}
+при \(i=1,\dots,M-2\) решаются только с использованием CPU.
+Непосредственно в данных подзадачах вычислений значений оптимизируемой функции не происходит: вычисление значения функции \(\phi_i(u_1,\dots,u_i)\) --- это решение задачи минимизации следующего уровня.
 Каждая подзадача решается в отдельном процессе; обмен данными осуществляется лишь между процессами-предками и процессами-потомками.
-Подзадача последнего (M–1)-го уровня
-
-отличается от всех предыдущих подзадач – в ней происходит вычисление значений оптимизируемой функции, т.к.
-.
-Подзадачи этого уровня решаются на сопроцессоре, и каждое ядро сопроцессора будет решать свою подзадачу (M–1)-го уровня в отдельном MPI-процессе.
-Самый простой вариант использования данной схемы будет соответствовать двухкомпонентному вектору распараллеливания  .
-Здесь   будет соответствовать числу MPI-процессов на CPU, а  числу MPI-процессов на Xeon Phi; тем самым общее количество процессов будет определяться как  .
+\par
+Подзадача последнего \((M–1)\)-го уровня 
+\begin{displaymath}
+\phi_{M-1}(u_1,\dots,u_{M-1})=\min_{u_{M}\in D_{M}}\phi_{M}(u_1,\dots,u_i,u_{M})
+\end{displaymath} 
+отличается от всех предыдущих подзадач --- в ней происходит вычисление значений оптимизируемой функции, т.к.
+\(\phi_M(u_1,\dots,u_M)=\phi(y_1,\dots,y_N)\).
+Подзадачи этого уровня решаются на сопроцессоре, и каждое ядро сопроцессора будет решать свою подзадачу \((M–1)\)-го уровня в отдельном MPI-процессе.
+\par
+Самый простой вариант использования данной схемы будет соответствовать двухкомпонентному вектору распараллеливания \(\pi=(\pi_1,\pi_2)\).
+Здесь \(\pi_1+1\) будет соответствовать числу MPI-процессов на CPU, а \(\pi_2\) --- числу MPI-процессов на Xeon Phi; тем самым общее количество процессов будет определяться как \(1+\pi_1+\pi_1\pi_2\).
+\par
 Отметим, что синхронный параллельный алгоритм глобальной оптимизации, описанный в п.3, в сочетании с блочной схемой редукции размерности обладает существенным недостатком, связанным с возможными простоями всех процессов, кроме корневого.
 Простои могут возникать в том случае, если часть потомков некоторого процесса закончили решение своих подзадач и отправили данные родителю раньше остальных.
-Для преодоления данной проблемы был применен предложенный в [12] асинхронный вариант параллельного алгоритма.
-Конкретные детали реализация блочной схемы в сочетании с асинхронным алгоритмом описаны в [12].
+Для преодоления данной проблемы был применен предложенный в \cite{examinArtcle} асинхронный вариант параллельного алгоритма.
+Конкретные детали реализация блочной схемы в сочетании с асинхронным алгоритмом описаны в \cite{examinArtcle}.

### 2016-02-15 2016/PAVT/numerical_experiments.tex
Update of numerical_experiments.tex

@@
 \par
-Ниже приведены результаты численного сравнения трех последовательных алгоритмов – DIRECT [7], DIRECTl [8] и алгоритма глобального поиска (АГП) (результаты рабо-ты первых двух алгоритмов приводятся по работе [9]).
-Численное сравнение проводилось на классах функций Simple и Hard размерности 4 и 5 из [9].
-Глобальный минимум \(y^*\) считался найденным, если алгоритм генерировал точку испытания yk в -окрестности глобального минимума, т.е.
-.
-При этом размер окрестности выбирался (в соответствии с [9]) как  , здесь N – размерность решаемой задачи, a и b – границы области поиска D, параметр =106 при N=4 и =107 при N=5.
-При использовании метода АГП для класса Simple выбирался параметр r=4.5, для класса Hard  r=5.6; параметр по-строения кривой Пеано был фиксированный m10.
-Максимально допустимое число итераций составляло Kmax = 1 000 000.
-В таблице 1 отражено среднее число итераций kav, которые выполнил метод при решении серии задач из данных классов.
+Ниже приведены результаты численного сравнения трех последовательных алгоритмов --- DIRECT \cite{zilinsk}, DIRECT\(l\)  \cite{sergKvaDiaPaper} и алгоритма глобального поиска (АГП) (результаты работы первых двух алгоритмов приводятся по работе \cite{jonesLipOpt}).
+Численное сравнение проводилось на классах функций Simple и Hard размерности 4 и 5 из \cite{jonesLipOpt}.
+Глобальный минимум \(y^*\) считался найденным, если алгоритм генерировал точку испытания \(y_k\) в \(\delta\)-окрестности глобального минимума, т.е.
+\(\Vert y_k-y^*\Vert\leqslant\delta\).
+При этом размер окрестности выбирался (в соответствии с \cite{jonesLipOpt}) как \(\delta=\Vert b-a\Vert \sqrt[N]{\Delta}\), здесь \(N\) --- размерность решаемой задачи, \(a\) и \(b\) --- границы области поиска \(D\), параметр \(\Delta=10^{-6}\) при \(N=4\) и \(\Delta=10^{-7}\) при \(N=5\).
+При использовании метода АГП для класса Simple выбирался параметр \(r=4.5\), для класса Hard --- \(r=5.6\); параметр построения кривой Пеано был фиксированный \(m=10\).
+Максимально допустимое число итераций составляло \(K_{max} = 1 000 000\).
+\par
+В таблице 1 отражено среднее число итераций \(k_{av}\), которые выполнил метод при решении серии задач из данных классов.
 Символ «>» отражает ситуацию, когда не все задачи класса были решены каким-либо методом.
-Это означает, что алгоритм был остановлен по причине достижения максимально допустимого числа итераций Kmax.
-В этом случае значение Kmax=1 000 000 использовалось при вычислении среднего значения числа итераций kav, что соответствует нижней оценке этого среднего значения.
+Это означает, что алгоритм был остановлен по причине достижения максимально допустимого числа итераций \(K_{max}\).
+В этом случае значение \(K_{max}=1 000 000\) использовалось при вычислении среднего значения числа итераций \(k_{av}\), что соответствует нижней оценке этого среднего значения.
 Количество нерешенных задач указано в скобках.
-Таблица 1. Среднее число итераций
-N	Класс функции	DIRECT	DIRECTl	АГП
-4	Simple	>47282 (4)	18983	11953
-	Hard	>95708 (7)	68754	25263
-5	Simple	>16057 (1)	16758	15920
-	Hard	>217215 (16)	>269064 (4)	>148342 (4)
-
-Как видно из таблицы 1, АГП превосходит методы DIRECT и DIRECTl на всех классах задач по среднему числу итераций.
-При этом в классе 5-Hard каждый из методов решил не все задачи: DIRECT не решил 16 задач, DIRECTl и АГП – по 4 задачи.
+\begin{table}
+    \centering
+    \begin{tabular}{|c|p{3cm}|c|c|c|}
+    \hline
+    \(N\) & Класс функций & DIRECT & DIRECT\(l\) & АГП \\ \hline
+    4 & Simple & >47282 (4) & 18983 & 11953\\ \cline{2-5}
+      & Hard & >95708 (7) & 68754 & 68754\\ \hline
+    5 & Simple & >16057 (1) & 16758 & 15920\\ \cline{2-5}
+      & Hard & >217215 (16) & >269064 (4) & >148342 (4)\\ \hline
+    \end{tabular}
+    \caption{Среднее число итераций}
+    \label{table:average_iters}
+\end{table}
+\par
+Как видно из таблицы \ref{table:average_iters}, АГП превосходит методы DIRECT и DIRECT\(l\) на всех классах задач по среднему числу итераций.
+При этом в классе 5-Hard каждый из методов решил не все задачи: DIRECT не решил 16 задач, DIRECT\(l\) и АГП – по 4 задачи.
 \subsection{Задачи с трудоемким критерием}
 Вначале приведем результаты экспериментов, которые соответствуют задачам с большим временем проведения одного поискового испытания.
-Оценим ускорение параллельного алгоритм глобального поиска (ПАГП), реализованного на CPU, в зависимости от использованного количества ядер p.
-В таблицах 2 и 3 приведено ускорение по времени S(p) и по итерациям s(p).
+Оценим ускорение параллельного алгоритм глобального поиска (ПАГП), реализованного на CPU, в зависимости от использованного количества ядер \(p\).
+В таблицах \ref{table:time_speedUp_cpu} и \ref{table:iterations_speedUp_cpu} приведено ускорение по времени \(S(p)\) и по итерациям \(s(p)\).
 Ускорение вычислено относительно однопоточного запуска.
-
-Таблица 2. Ускорение по времени S(p) на CPU
-p	N4	N5
-	Simple	Hard	Simple	Hard
-2	2.45	2.20	1.15	1.32
-4	4.66	3.90	2.82	2.59
-8	7.13	7.35	3.47	5.34
-
-Таблица 3. Ускорение по итерациям s(p) на CPU
-p	N4	N5
-	Simple	Hard	Simple	Hard
-2	2.51	2.26	1.19	1.36
-4	5.04	4.23	3.06	2.86
-8	8.58	8.79	4.22	6.56
-
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
+    & Simple & Hard & Simple & Hard \\ \hline
+    2 & 2.45 & 2.20 & 1.15 & 1.32\\ \hline
+	4 & 4.66 & 3.90 & 2.82 & 2.59\\ \hline
+	8 & 7.13 & 7.35 & 3.47 & 5.34\\ \hline
+    \end{tabular}
+    \caption{Ускорение по времени \(S(p)\) на CPU}
+    \label{table:time_speedUp_cpu}
+\end{table}
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
+    & Simple & Hard & Simple & Hard \\ \hline
+	2 & 2.51 & 2.26 & 1.19 & 1.36\\ \hline
+	4 & 5.04 & 4.23 & 3.06 & 2.86\\ \hline
+	8 & 8.58 & 8.79 & 4.22 & 6.56\\ \hline
+    \end{tabular}
+    \caption{Ускорение по итерациям \(s(p)\) на CPU}
+    \label{table:iterations_speedUp_cpu}
+\end{table}
+\par
 Результаты экспериментов показывают значительное ускорение ПАГП при использовании CPU. Лучшие результаты получены при использовании всех ядер процессора.
+\par
 Далее изложим результаты экспериментов, проведенных с использованием Intel Xeon Phi.
@@
 Варьировалось количество потоков на сопроцессоре, все остальные параметры совпадают с предыдущими запусками.
-Приведено (таблицы 4 и 5) ускорение относительно восьмипоточного запуска на центральном процессоре.
-Таблица 4. Ускорение по времени S(p) на Phi
-p	N4	N5
-	Simple	Hard	Simple	Hard
-60	0.54	1.02	1.07	1.61
-120	0.55	1.17	1.05	2.61
-240	0.51	1.06	1.07	4.17
-Таблица 5. Ускорение по итерациям s(p) на Phi
-p	N4	N5
-	Simple	Hard	Simple	Hard
-60	8.13	7.32	9.87	6.55
-120	16.33	15.82	15.15	17.31
-240	33.07	27.79	38.80	59.31
-
-Результаты экспериментов показывают, что только на классе Simple при N=4 реализация на Xeon Phi медленнее, чем CPU реализация, на классе Hard при N=4 и на классе Sim-ple при N=5 версии для Xeon Phi и CPU демонстрирует примерно равную эффективность, а на классе Hard при N=5 версия для Xeon Phi существенно превосходит CPU-реализацию.
-Наибольшее преимущество реализация для Xeon Phi показывает при 120 потоках для че-тырехмерных задач и при 240 потоках для пятимерных.
+Приведено (таблицы \ref{table:time_speedUp_phi} и \ref{table:iterations_speedUp_phi}) ускорение относительно восьмипоточного запуска на центральном процессоре.
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
+    & Simple & Hard & Simple & Hard \\ \hline
+	60 & 0.54 & 1.02 & 1.07 & 1.61\\ \hline
+	120 & 0.55 & 1.17 & 1.05 & 2.61\\ \hline
+	240 & 0.51 & 1.06 & 1.07 & 4.17\\ \hline
+    \end{tabular}
+    \caption{Ускорение по времени \(S(p)\) на Phi}
+    \label{table:time_speedUp_phi}
+\end{table}
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
+    & Simple & Hard & Simple & Hard \\ \hline
+	60 & 8.13 & 7.32 & 9.87 & 6.55 \\ \hline
+	120 & 16.33 & 15.82 & 15.15 & 17.31 \\ \hline
+	240 & 33.07 & 27.79 & 38.80 & 59.31 \\ \hline
+    \end{tabular}
+    \caption{Ускорение по итерациям \(s(p)\) на Phi}
+    \label{table:iterations_speedUp_phi}
+\end{table}
+\par
+Результаты экспериментов показывают, что только на классе Simple при \(N=4\) реализация на Xeon Phi медленнее, чем CPU реализация, на классе Hard при \(N=4\) и на классе Simple при \(N=5\) версии для Xeon Phi и CPU демонстрирует примерно равную эффективность, а на классе Hard при \(N=5\) версия для Xeon Phi существенно превосходит CPU-реализацию.
+Наибольшее преимущество реализация для Xeon Phi показывает при 120 потоках для четырехмерных задач и при 240 потоках для пятимерных.
 При этом на всех классах наблюдается значительное ускорение по итерациям, а также практически линейная масштабируемость от числа потоков.
+\par
 Далее рассмотрим эксперименты с запуском сопроцессора в режиме MPI.
 Число уровней вложенности разбиваемой задачи равно двум.
-Число процессов, запущенных на со-процессоре, – 30 (таблицы 6 и 7) и 60(таблицы 8 и 9), что соответствует ситуации, когда у корня дерева имеется соответственно 30 и 60 потомков.
+Число процессов, запущенных на сопроцессоре, –-- 30 (таблицы \ref{table:time_speedUp_phi_30mpi} и \ref{table:iterations_speedUp_phi_30mpi}) и 60(таблицы \ref{table:time_speedUp_phi_60mpi} и \ref{table:iterations_speedUp_phi_60mpi}), что соответствует ситуации, когда у корня дерева имеется соответственно 30 и 60 потомков.
 Процессы, работающие на сопроцессоре, использовали OpenMP для параллельного вычисления значений функции.
 Ускорение приведено относительно восьмипоточного запуска на CPU.
-Таблица 6. Ускорение по времени S(p) на Phi, 30 MPI процессов
-p	N4	N5
-	Simple	Hard	Simple	Hard
-4	3,18	5,14	5,15	14,65
-8	3,53	1,49	3,35	10,49
-16	1,09	2,77	4,61	14,36
-Таблица 7. Ускорение по итерациям s(p) на Phi, 30 MPI процессов
-p	N4	N5
-	Simple	Hard	Simple	Hard
-4	11,19	19,19	9,27	27,38
-8	13,97	3,63	8,24	23,85
-16	1,29	3,56	8,28	22,12
-Таблица 8. Ускорение по времени S(p) на Phi, 60 MPI процессов
-p	N4	N5
-	Simple	Hard	Simple	Hard
-4	2,76	3,20	4,19	16,56
-8	3,05	7,73	4,08	14,87
-16	0,89	5,21	2,72	13,05
-Таблица 9. Ускорение по итерациям s(p) на Phi, 60 MPI процессов
-p	N4	N5
-	Simple	Hard	Simple	Hard
-4	10,36	3,34	6,01	26,04
-8	9,04	28,81	7,68	25,02
-16	1,27	28,66	7,19	30,38
-
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
+    & Simple & Hard & Simple & Hard \\ \hline
+	4 & 3,18 & 5,14 & 5,15 & 14,65 \\ \hline
+	8 & 3,53 & 1,49 & 3,35 & 10,49 \\ \hline
+	16 & 1,09 & 2,77 & 4,61 & 14,36 \\ \hline
+    \end{tabular}
+    \caption{Ускорение по времени \(S(p)\) на Phi, 30 MPI процессов}
+    \label{table:time_speedUp_phi_30mpi}
+\end{table}
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
+	4 & 11,19 & 19,19 & 9,27 & 27,38 \\ \hline
+	8 & 13,97 & 3,63 & 8,24 & 23,85 \\ \hline
+	16 & 1,29 & 3,56 & 8,28 & 22,12 \\ \hline
+	\end{tabular}
+    \caption{Ускорение по итерациям \(s(p)\) на Phi, 30 MPI процессов}
+    \label{table:iterations_speedUp_phi_30mpi}
+\end{table}
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
+    & Simple & Hard & Simple & Hard \\ \hline
+	4 & 2,76 & 3,20 & 4,19 & 16,56 \\ \hline
+	8 & 3,05 & 7,73 & 4,08 & 14,87 \\ \hline
+	16 & 0,89 & 5,21 & 2,72 & 13,05 \\ \hline
+    \end{tabular}
+    \caption{Ускорение по времени \(S(p)\) на Phi, 60 MPI процессов}
+    \label{table:time_speedUp_phi_60mpi}
+\end{table}
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
+	4 & 10,36 & 3,34 & 6,01 & 26,04 \\ \hline
+	8 & 9,04 & 28,81 & 7,68 & 25,02 \\ \hline
+	16 & 1,27 & 28,66 & 7,19 & 30,38 \\ \hline
+	\end{tabular}
+    \caption{Ускорение по итерациям \(s(p)\) на Phi, 60 MPI процессов}
+    \label{table:iterations_speedUp_phi_60mpi}
+\end{table}
+\par
 Результаты экспериментов показывают, что запуск в режиме MPI значительно превосходит Offload режим по времени работы, однако проигрывает по числу итераций.
 Лучшие результаты для простого класса задач наблюдаются при использовании 30 процессов на сопроцессоре, а для сложного класса, при 60 процессах.
-Таким образом, можно сделать вывод, что задачи с большим временем работы, для которых важно число итераций мето-да оптимизации, лучше использовать запуск в режиме Offload, а для более быстро решаемых задач, лучше подходит режим MPI.
+Таким образом, можно сделать вывод, что задачи с большим временем работы, для которых важно число итераций метода оптимизации, лучше использовать запуск в режиме Offload, а для более быстро решаемых задач, лучше подходит режим MPI.
 \subsection{Задачи с легко вычисляемым критерием}
-Рассмотрим теперь решение задач, в которых поисковые испытания проводятся быстро. Вначале приведем результаты OpenMP версии. В таблице 10 приведено ускорение относительно однопоточного запуска.
-Таблица 10. Ускорение по времени S(p) на CPU
-p	N4	N5
-	Simple	Hard	Simple	Hard
-2	1,09	1,09	0,56	0,57
-4	1,10	1,06	0,74	0,66
-8	0,90	1,11	0,45	1,77
-16	0,66	0,92	0,76	0,66
-32	0,29	0,38	0,22	0,55
-
+Рассмотрим теперь решение задач, в которых поисковые испытания проводятся быстро.
+Вначале приведем результаты OpenMP версии.
+В таблице \ref{table:time_speedUp_cpu_2} приведено ускорение относительно однопоточного запуска.
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
+    & Simple & Hard & Simple & Hard \\ \hline
+	2 & 1,09 & 1,09 & 0,56 & 0,57 \\ \hline
+	4 & 1,10 & 1,06 & 0,74 & 0,66 \\ \hline
+	8 & 0,90 & 1,11 & 0,45 & 1,77 \\ \hline
+	16 & 0,66 & 0,92 & 0,76 & 0,66 \\ \hline
+	32 & 0,29 & 0,38 & 0,22 & 0,55 \\ \hline
+    \end{tabular}
+    \caption{Ускорение по времени \(S(p)\) на CPU}
+    \label{table:time_speedUp_cpu_2}
+\end{table}
+\par
 Эксперименты показывают незначительное ускорение, а во многих случаях – и замедление при вычислениях только на центральном процессоре.
+\par
 Далее рассмотрим ускорение при использовании Xeon Phi в режиме MPI.
 Результаты приведены относительно однопоточного запуска на CPU.
-В таблице 11 приведено ускорение по времени S(p) для запуска 30 процессов на Xeon Phi, в таблице 12 для 60 процессов.
-Таблица 11. Ускорение по времени S(p) на Phi, 30 MPI процессов
-p	N4	N5
-	Simple	Hard	Simple	Hard
-1	0,35	2,28	0,37	1,92
-2	0,94	1,95	0,39	1,23
-4	1,23	0,43	0,54	2,63
-8	0,85	0,56	0,40	1,53
-
-Таблица 12. Ускорение по времени S(p) на Phi, 60 MPI процессов
-p	N4	N5
-	Simple	Hard	Simple	Hard
-1	0,59	2,56	1,22	3,30
-2	0,89	0,87	1,29	2,76
-4	1,09	1,03	0,38	5,32
-8	0,70	0,78	0,69	2,19
-
+В таблице \ref{table:time_speedUp_phi_30mpi_easy} приведено ускорение по времени \(S(p)\) для запуска 30 процессов на Xeon Phi, в таблице \ref{table:time_speedUp_phi_60mpi_easy} для 60 процессов.
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
+    & Simple & Hard & Simple & Hard \\ \hline
+	1 & 0,35 & 2,28 & 0,37 & 1,92 \\ \hline
+	2 & 0,94 & 1,95 & 0,39 & 1,23 \\ \hline
+	4 & 1,23 & 0,43 & 0,54 & 2,63 \\ \hline
+	8 & 0,85 & 0,56 & 0,40 & 1,53 \\ \hline
+	\end{tabular}
+    \caption{Ускорение по времени \(S(p)\) на Phi, 30 MPI процессов}
+    \label{table:time_speedUp_phi_30mpi_easy}
+\end{table}
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
+    & Simple & Hard & Simple & Hard \\ \hline
+	1 & 0,59 & 2,56 & 1,22 & 3,30 \\ \hline
+	2 & 0,89 & 0,87 & 1,29 & 2,76 \\ \hline
+	4 & 1,09 & 1,03 & 0,38 & 5,32 \\ \hline
+	8 & 0,70 & 0,78 & 0,69 & 2,19 \\ \hline
+	\end{tabular}
+    \caption{Ускорение по времени \(S(p)\) на Phi, 60 MPI процессов}
+    \label{table:time_speedUp_phi_60mpi_easy}
+\end{table}
+\par
 Результаты экспериментов показывают ускорение на Xeon Phi, превосходящее ускорение на центральном процессоре.
 Лучшие результаты получены при использовании 30 процессов на Xeon Phi только для простых четырехмерных задач, все остальные задачи лучше всего решаются при использовании 60 процессов.
+\par
 Теперь перейдем к эксперименту с шестимерными задачами простого класса.
-При решении шестимерной задачи с использованием одного процесса (вложенные подзадачи отсутствуют) и сохранением точности построения развертки m=10 требуется использовать числа с плавающей запятой расширенной точности, что ведет к значительному увеличению времени работы алгоритма.
+При решении шестимерной задачи с использованием одного процесса (вложенные подзадачи отсутствуют) и сохранением точности построения развертки \(m=10\) требуется использовать числа с плавающей запятой расширенной точности, что ведет к значительному увеличению времени работы алгоритма.
 При этом в случае использования хотя бы одного уровня вложенных подзадач, задачи на каждом уровне будут решаться без подключения расширенной точности, что обеспечивает дополнительное преимущество блочной многошаговой схемы.
-Рассмотрим результаты эксперимента с использованием Xeon Phi в режиме MPI, таблица 13.
+\par
+Рассмотрим результаты эксперимента с использованием Xeon Phi в режиме MPI, таблица \ref{table:time_speedUp_phi_2}.
 Число уровней вложенности разбиваемой задачи равно двум.
-Приведены результаты экспериментов для разбиений 3:3 (три переменных на первом уровне, три – на вто-ром) и 4:2 (4 переменных на первом уровне, две – на втором).
+Приведены результаты экспериментов для разбиений 3:3 (три переменных на первом уровне, три – на втором) и 4:2 (4 переменных на первом уровне, две – на втором).
 Число процессов, запущенных на сопроцессоре равно 30 и 60, что соответствует тому, что у корня дерева имеется соответственно 30 и 60 потомков.
@@
 Ускорение приведено относительно однопоточного запуска на CPU.
-Таблица 13. Ускорение по времени S(p) на Phi
-p	30 процессов	60 процессов
-	3:3	4:2	3:3	4:2
-2	7,30	9,87	17,68	18,21
-4	5,99	8,18	6,60	13,57
-8	3,14	12,74	5,92	25,58
-16	3,54	7,61	4,87	11,33
-32	3,41	6,85	3,73	6,50
-
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \(p\) & \multicolumn{2}{|c|}{30 процессов} & \multicolumn{2}{|c|}{60 процессов}\\ \cline{2-5}
+	& 3:3 & 4:2 & 3:3 & 4:2 \\ \hline
+	2 & 7,30 & 9,87 & 17,68 & 18,21 \\ \hline
+	4 & 5,99 & 8,18 & 6,60 & 13,57 \\ \hline
+	8 & 3,14 & 12,74 & 5,92 & 25,58 \\ \hline
+	16 & 3,54 & 7,61 & 4,87 & 11,33 \\ \hline
+	32 & 3,41 & 6,85 & 3,73 & 6,50 \\ \hline
+\end{tabular}
+    \caption{Ускорение по времени \(S(p)\) на Phi}
+    \label{table:time_speedUp_phi_2}
+\end{table}
+\par
 Результаты экспериментов показывают значительное ускорение запуска на сопроцессоре по сравнению с запуском только на центральном процессоре.

### 2016-02-16 2016/PAVT/main_eng.tex
Add english abstract and refs

\documentclass[a4paper]{article}
\usepackage{fontspec}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[a4paper]{geometry}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\setmainfont{CMU Serif}
\setsansfont{CMU Sans Serif}

\bibliographystyle{unsrt}
\fontsize{10}{12pt}\selectfont

\begin{document}

\title{Implementation of the parallel asynchronous global search algorithm on Intel Xeon Phi\
\footnote{
The paper was supported by the Russian Science Foundation, project No 15-11-30022 “Global optimization, supercomputing  computations, and applications”}}
\author{K.A. Barkalov, I.G. Lebedev, V.V. Sovrasov, A.V. Sysoyev}
\date{}
\maketitle

\begin{abstract}
The paper considers parallel algorithm for solving multiextremal optimization problems. The issues of implementation of the algorithm on state-of-the-art computing systems using Intel Xeon Phi coprocessor are examined. Two approaches for algorithm parallelization, which take into account information about laboriousness of the objective function computing, are considered. Speed up of the algorithm using Xeon Phi compared to the algorithm using CPU only is experimentally confirmed. Computational experiments are carried out on Lobachevsky supercomputer.
\par
\textit{Keywords:}global optimization, multiextremal functions, dimension reduction, parallel computing, Intel Xeon Phi.
\end{abstract}
\nocite{*}
\bibliography{refs_eng}
\end{document}

### 2016-02-16 2016/PAVT/numerical_experiments.tex
Add english abstract and refs

@@
 \par
-Ниже приведены результаты численного сравнения трех последовательных алгоритмов --- DIRECT \cite{zilinsk}, DIRECT\(l\)  \cite{sergKvaDiaPaper} и алгоритма глобального поиска (АГП) (результаты работы первых двух алгоритмов приводятся по работе \cite{jonesLipOpt}).
-Численное сравнение проводилось на классах функций Simple и Hard размерности 4 и 5 из \cite{jonesLipOpt}.
+Ниже приведены результаты численного сравнения трех последовательных алгоритмов --- DIRECT \cite{jonesLipOpt}, DIRECT\(l\)  \cite{gablonskyDirect} и алгоритма глобального поиска (АГП) (результаты работы первых двух алгоритмов приводятся по работе \cite{sergKvaDiaPaper}).
+Численное сравнение проводилось на классах функций Simple и Hard размерности 4 и 5 из \cite{sergKvaDiaPaper}.
 Глобальный минимум \(y^*\) считался найденным, если алгоритм генерировал точку испытания \(y_k\) в \(\delta\)-окрестности глобального минимума, т.е.
 \(\Vert y_k-y^*\Vert\leqslant\delta\).
-При этом размер окрестности выбирался (в соответствии с \cite{jonesLipOpt}) как \(\delta=\Vert b-a\Vert \sqrt[N]{\Delta}\), здесь \(N\) --- размерность решаемой задачи, \(a\) и \(b\) --- границы области поиска \(D\), параметр \(\Delta=10^{-6}\) при \(N=4\) и \(\Delta=10^{-7}\) при \(N=5\).
+При этом размер окрестности выбирался (в соответствии с \cite{sergKvaDiaPaper}) как \(\delta=\Vert b-a\Vert \sqrt[N]{\Delta}\), здесь \(N\) --- размерность решаемой задачи, \(a\) и \(b\) --- границы области поиска \(D\), параметр \(\Delta=10^{-6}\) при \(N=4\) и \(\Delta=10^{-7}\) при \(N=5\).
 При использовании метода АГП для класса Simple выбирался параметр \(r=4.5\), для класса Hard --- \(r=5.6\); параметр построения кривой Пеано был фиксированный \(m=10\).

### 2016-03-04 2016/PAVT/main_eng.tex
Move bibliography to main_single.tex main_single.tex and main_eng.tex are ready for sending to journal

@@
 \nocite{*}
-\bibliography{refs_eng}
+
+\begin{thebibliography}{10}
+
+\bibitem{evtushenko}
+Evtushenko Yu.G.
+\newblock {\em Metody resheniya ekstremal'nykh zadach i ikh primenenie v
+  sistemakh optimizatsii [Methods of Solving Extremal Problems and Their
+  Application in Optimization Systems]}.
+\newblock Moscow: Science, 1982. 432 p.
+
+\bibitem{pinter}
+Pinter J.
+\newblock {\em Global optimization in action (Continuous and Lipschitz
+  optimization: algorithms, imple-mentations and applications)}.
+\newblock Kluwer Academic Publishers, 1996, 507 p.
+
+\bibitem{sergKvaDiag}
+Kvasov~D.E. Sergeyev~Ya.D.
+\newblock {\em Diagonal'nye metody global'noy optimizatsii [Diagonal global
+  optimization methods]}.
+\newblock Moscow: Fizmatlit, 2008. 352 p.
+
+\bibitem{strSergOptBook}
+Sergeyev~Ya.D. Strongin~R.G.
+\newblock {\em Global Optimization with Non-convex Constraints. Sequential and
+  Parallel Algorithms}.
+\newblock Kluwer Academic Publishers, 2000, 704p.
+
+\bibitem{strGergrParOptBook}
+Strongin R.G. Gergel V.P. Grishagin V.A.~Barkalov K.A.
+\newblock {\em Parallel'nye vychisleniya v zadachakh global'noy optimizatsii
+  [Parallel computing in global optimization problems]}.
+\newblock Moscow: MSU Publishing, 2013. 280p.
+
+\bibitem{sergKvaDiaPaper}
+Sergeyev Ya.D.~Kvasov D.E.
+\newblock Global search based on efficient diagonal partitions and a set of
+  lipschitz constants.
+\newblock {\em SIAM Journal on Optimization}, 16(3):910–937, 2006.
+
+\bibitem{zilinsk}
+Žilinskas J.
+\newblock Branch and bound with simplicial partitions for global optimization.
+\newblock {\em Mathematical Modelling and Analysis}, 13(1):145–159, 2008.
+
+\bibitem{gorodGrishOptBook}
+Gorodetskiy S.Yu.~Grishagin V.A.
+\newblock {\em Nelineynoe programmirovanie i mnogoekstremal'naya optimizatsiya
+  [Nonlinear pro-gramming and multiextremal optimization]}.
+\newblock N. Novgorod: UNN Publishing, 2007.
+
+\bibitem{examinArtcle}
+Sysoyev A.V. Barkalov K.A. Gergel V.P.~Lebedev I.G.
+\newblock Mpi-realizatsiya blochnoy mnogoshagovoy skhemy parallel'nogo
+  resheniya zadach global'noy optimizatsii [mpi-implementation of the nested
+  block scheme for parallel solving of global optimization problems ].
+\newblock In {\em Superkomp'yuternye dni v Rossii: Trudy mezhdunarodnoy
+  konferentsii (28-29 sen-tyabrya 2015 g., g. Moskva) [Russian supercomputing
+  days: proceedings of the interna-tional conference (Moscow, September 28-29,
+  2015)]}, pages 411--419. Moscow: MSU Publishing, 2015.
+
+\bibitem{gklsPaper}
+Gaviano M. Lera D. Kvasov D.E.~Sergeyev Ya.D.
+\newblock Software for generation of classes of test functions with known local
+  and global minima for global optimization.
+\newblock {\em ACM Transactions on Mathematical Software}, 29:469--480, 2003.
+
+\bibitem{jonesLipOpt}
+Jones D.R. Perttunen C.D.~Stuckman B.E.
+\newblock Lipschitzian optimization without the lipschitz constant.
+\newblock {\em Journal of Optimization Theory and Applications},
+  79(1):157–181, 1993.
+
+\bibitem{gablonskyDirect}
+Kelley~C.T. Gablonsky~J.M.
+\newblock A locally-biased form of the direct algorithm.
+\newblock {\em Journal of Global Optimization}, 21(1):27--37, 2001.
+
+\end{thebibliography}
+
 \end{document}

### 2016-03-04 2016/PAVT/main_single.tex
Move bibliography to main_single.tex main_single.tex and main_eng.tex are ready for sending to journal

\documentclass[a4paper]{article}
\usepackage{fontspec}
\usepackage[russian]{babel}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[a4paper]{geometry}
\usepackage{indentfirst}
\usepackage{multirow}
\usepackage{graphicx}

\setmainfont{CMU Serif}
\setsansfont{CMU Sans Serif}

\bibliographystyle{unsrt}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\sign}{sign}
\fontsize{10}{12pt}\selectfont
\begin{document}

\title{Реализация параллельного алгоритма поиска глобального экстремума функции на Intel Xeon Phi\
\footnote{
Исследование выполнено при финансовой поддержке РФФИ в рамках научного проекта № 16-31-00244 мол\_а «Параллельные методы решения вычислительно трудоемких задач глобальной оптимизации на гибридных кластерных системах»}}
\author{К.А. Баркалов, И.Г. Лебедев, В.В. Соврасов, А.В. Сысоев}
\date{}
\maketitle

\begin{abstract}
Рассмотрен параллельный алгоритм решения задач многоэкстремальной оптимизации. Описывается реализация алгоритма на современных вычислительных системах с использованием сопроцессора Xeon Phi. Рассматриваются два подхода к распараллеливанию алгоритма, учитывающие информацию о трудоемкости вычисления значений оптимизируемой функции. Приводятся результаты вычислительных экспериментов, полученные на суперкомпьютере «Лобачевский». Демонстрируется, что реализация для Xeon Phi опережает версию для CPU. Результаты подтверждают ускорение алгоритма с использованием Xeon Phi по сравнению с алгоритмом, реализованным только на CPU.
\par
\textit{Ключевые слова:} глобальная оптимизация, многоэкстремальные функции, редукция размерности, параллельные алгоритмы, Intel Xeon Phi.
\end{abstract}

\section{Введение}
Рассматриваются задачи многоэкстремальной оптимизации и параллельные методы их решения. Важной особенностью указанных задач является тот факт, что глобальный экстремум есть интегральная характеристика задачи, таким образом, его отыскание связано с построением покрытия области поиска и вычислением значений оптимизируемой функции во всех точках этого покрытия. На сложность решения задач рассматриваемого класса решающее влияние оказывает размерность: вычислительные затраты растут экспоненциально при ее увеличении. Использование простейших способов решения (таких, как перебор на равномерной сетке) является неприемлемым. Требуется применение более экономных методов, которые порождают в области поиска существенно неравномерную сетку, более плотную в окрестности глобального минимума и разреженную вдали от него \cite{evtushenko}\cite{pinter}\cite{sergKvaDiag}. Данная статья продолжает развитие информационно-статистического подхода к построению параллельных алгоритмов глобальной оптимизации, предложенного в ННГУ им. Н.И. Лобачевского, который описан в монографиях \cite{strSergOptBook}\cite{strGergrParOptBook}.
\par
В рамках обсуждаемого подхода решение многомерных задач сводится к решению набора связанных подзадач меньшей размерности. Соответствующая редукция основана на использовании разверток единичного отрезка вещественной оси на гиперкуб. Роль таких разверток играют непрерывные однозначные отображения типа кривой Пеано, называемые также кривыми, заполняющими пространство. Еще одним используемым механизмом снижения размерности решаемой задачи является схема вложенной (рекурсивной) оптимизации. Численные методы, позволяющие эффективно использовать аппарат таких отображений, детально разработаны и обоснованы в \cite{strSergOptBook}\cite{strGergrParOptBook}.
\par
Алгоритмы, развиваемые в рамках информационно-статистического подхода, основаны на предположении липшицевости оптимизируемого критерия, что является типичными для других методов (см., например, \cite{pinter}\cite{sergKvaDiag}). Предположение такого рода выполняется для многих прикладных задач, поскольку относительные вариации функций, характеризующих моделируемую систему, обычно не превышают некоторый порог, определяемый ограниченной энергией изменений в системе.
\par
Использование современных параллельных вычислительных систем расширяет сферу применения методов глобальной оптимизации и, в то же время, ставит задачу эффективно-го распараллеливания процесса поиска. Именно поэтому разработка эффективных параллельных методов для численного решения задач многоэкстремальной оптимизации и со-здание на их основе программных средств для современных вычислительных систем является актуальной задачей. Особый интерес представляет разработка схем распараллеливания, позволяющих эффективно использовать ускорители вычислений, такие, как сопроцессор Intel Xeon Phi.
\par
В рамках проводимого исследования мы будем предполагать, что время проведения одного испытания (вычисления значения функции в области поиска) может значительно отличаться в различных решаемых задачах. Это определяет разработку разных подходов к распараллеливанию в задачах с «легкими» и «сложными» критериями. В статье приведено описание универсального подхода к распараллеливанию алгоритма глобального поиска, который охватывает оба этих случая. Указанный подход реализован в разработанной в ННГУ им. Н.И. Лобачевского параллельной программной системе решения задач глобальной оптимизации.

\section{Параллельный алгоритм глобального поиска}
Рассмотрим задачу поиска глобального минимума \(N\)-мерной функции \(\phi(y)\) в гиперинтервале \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\). Будем предполагать, что функция удовлетворяет условию Липшица с априори неизвестной константой \(L\).
\begin{equation}
\label{task}
\phi(y^*)=\min\{\phi(y):y\in D\}
\end{equation}
\begin{equation}
\label{lip}
|\phi(y_1)-\phi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
\end{equation}
\par
Существует ряд способов адаптации эффективных одномерных алгоритмов для решения многомерных задач, см., например, методы диагонального \cite{sergKvaDiaPaper} или симплексного \cite{zilinsk} разбиения области поиска. В данной работе мы будем использовать подход, основанный на идее редукции размерности с помощью кривой Пеано \(y(x)\), непрерывно и однозначно отображающей отрезок вещественной оси \([0,1]\) на \(n\)-мерный куб 
\begin{displaymath}
\label{cube}
\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
\end{displaymath}
\par
Вопросы численного построения отображений типа кривой Пеано и соответствующая теория подробно рассмотрены в \cite{strSergOptBook}. Здесь же отметим, что численно построенная развертка является приближением к теоретической кривой Пеано с точностью порядка \(2^{-m}\), где \(m\) –-- параметр построения развертки. Использование подобного рода отображений позволяет свести многомерную задачу к одномерной задаче
\begin{displaymath}
\label{oneDimTask}
\phi(y^*)=\phi(y(x^*))=\min\{\phi(y(x)):x\in [0,1]\}
\end{displaymath}
\par
Важным свойством является сохранение ограниченности относительных разностей функции: если функция \(\phi(y)\) в области \(D
\) удовлетворяла условию Липшица, то функция \(\phi(y(x))\) на интервале \([0,1]\) будет удовлетворять равномерному условию Гельдера
\begin{displaymath}
\label{holder}
|\phi(y(x_1))-\phi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
\end{displaymath}
где константа Гельдера \(H\) связана с константой Липшица \(L\) соотношением 
\begin{displaymath}
H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}
\end{displaymath}
\par
Поэтому, не ограничивая общности, можно рассматривать минимизацию одномерной функции  \(f(x)=\phi(y(x)), x\in[0,1]\), удовлетворяющей условию Гельдера.
\par
Рассматриваемый алгоритм решения данной задачи предполагает построение последовательности точек \(x_k\), в которых вычисляются значения минимизируемой функции \(z_k = f(x_k)\). Процесс вычисления значения функции (включающий в себя построение образа \(y_k=y(x_k))\) будем называть испытанием, а пару \((x_k,z_k)\) --- результатом испытания. Множество пар \(\{(x_k,z_k)\}, 1\leqslant k\leqslant n\) составляют поисковую информацию, накопленную методом после проведения \(n\) шагов. В нашем распоряжении имеется \(p\geqslant 1\) вычислительных элементов и в рамках одной итерации метода мы будем проводить \(p\) испытаний одновременно. Обозначим \(k(n)\) общее число испытаний, выполненных после \(n\) параллельных итераций.
\par
На первой итерации метода испытание проводится в произвольной внутренней точке \(x_1\) интервала \([0,1]\). Пусть выполнено \(n\geqslant 1\) итераций метода, в процессе которых были проведены испытания в \(k = k(n)\) точках \(x_i, 1\leqslant i\leqslant k\). Тогда точки \(x^{k+1},\dotsc,x^{k+p}\) поисковых испытаний следующей \((n+1)\)-ой итерации определяются в соответствии с правилами:
\par
Шаг 1. Перенумеровать точки множества \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), которое включает в себя граничные точки интервала \([0,1]\), а также точки предшествующих испытаний, нижними индексами в порядке увеличения значений координаты, т.е.
\begin{displaymath}
0=x_0<x_1<\dotsc<x_{k+1}=1
\end{displaymath}
\par
Шаг 2. Полагая \(z_i=f(x_i),1\leqslant i\leqslant k\), вычислить величины 
\begin{equation}
\label{step2}
\mu=\max_{1\leqslant i\leqslant k}\dfrac{|z_i-z_{i-1}|}{\Delta_i},
\begin{matrix}
    M = 
    \left\{
    \begin{matrix}
    r\mu,\mu>0 \\
    1,\mu=0
    \end{matrix} \right.
    \end{matrix}
\end{equation}
где \(r\) является заданным параметром метода, а \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
\par
Шаг 3. Для каждого интервала \((x_{i-1},x_i),1\leqslant i\leqslant k+1\), вычислить характеристику в соответствии с формулами
\begin{equation}
\label{step3_1}
R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M}
\end{equation}
\begin{equation}
\label{step3_2}
R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1
\end{equation}
\par
Шаг 4. Характеристики \(R(i),1\leqslant i\leqslant k+1\), упорядочить в порядке убывания 
\begin{equation}
\label{step4}
R(t_1)\geqslant R(t_2)\geqslant \dots \geqslant R(t_{k})\geqslant R(t_{k+1})
\end{equation}
и выбрать \(p\) наибольших характеристик с номерами интервалов \(t_j,1\leqslant j\leqslant p\).
\par
Шаг 5. Провести новые испытания в точках \(x_{k+j},1\leqslant j\leqslant p\), вычисленных по формулам
\begin{displaymath}
x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},t_j=1,t_j=k+1
\end{displaymath}
\begin{equation}
\label{step5}
x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-\sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{\mu}\right]^N,1<t_j<k+1
\end{equation}
\par
Алгоритм прекращает работу, если выполняется условие \(\Delta_{t_j}\leqslant \varepsilon\) хотя бы для одного номера \(t_j,1\leqslant j\leqslant p\); здесь \(\varepsilon>0\) есть заданная точность. В качестве оценки глобально-оптимального решения задачи  выбираются значения
\begin{equation}
f_k^*=\min_{1\leqslant i \leqslant k}f(x_i), x_k^*=\argmin_{1\leqslant i \leqslant k}f(x_i)
\end{equation}
\par
Теоретическое обоснование данного способа организации параллельных вычислений изложено в \cite{strGergrParOptBook}.

\section{Обобщенная схема редукция размерности}
Одним из подходов к решению многомерных задач глобальной оптимизации является сведение их к одномерным и использование эффективных одномерных алгоритмов глобального поиска к редуцированной задаче. В предыдущем разделе была изложена идея редукции размерности с использованием кривых Пеано. Ниже излагается обобщенный способ редукции размерности, комбинирующий использование разверток и схему вложенной (рекурсивной) оптимизации.
\subsection{Рекурсивная схема редукции размерности}
Схема рекурсивной оптимизации основана на известном \cite{gorodGrishOptBook} соотношении
\begin{equation}
\label{nestedScheme}
\min{\phi(y):y\in D}=\min_{a_1\leqslant y_1\leqslant b_1}\min_{a_2\leqslant y_2\leqslant b_2}\dots\min_{a_1\leqslant y_N\leqslant b_N}\phi(y)
\end{equation}
которое позволяет заменить решение многомерной задачи (\ref{task}) решением семейства одно-мерных подзадач, рекурсивно связанных между собой.
\par
Введем в рассмотрение множество функций 
\begin{equation}
\label{phiN}
\phi_N(y_1,\dots,y_N)=\phi(y_1,\dots,y_N)
\end{equation}
\begin{equation}
\phi_i(y_1,\dots,y_i)=\min_{a_{i+1}\leqslant y_{i+1} \leqslant b_{i+1}}\phi_{i+1}(y_1,\dots,y_i,y_{i+1}),1\leqslant i\leqslant N-1
\end{equation}
\par
Тогда, в соответствии с соотношением (\ref{nestedScheme}), решение исходной задачи  сводится к решению одномерной задачи
\begin{equation}
\label{phiFirst}
\phi_1(y_1^*)=\min\{\phi_1(y_1),y_1\in [a,b]\}
\end{equation}
\par
Однако при этом каждое вычисление значения одномерной функции \(\phi_1(y_1)\) в некоторой фиксированной точке предполагает решение одномерной задачи минимизации \(\phi_2(y_1,y_2^*)=\min\{\phi(y_1,y_2):y_2\in [a_2,b_2]\}\), и так далее до вычисления \(\phi_N\) согласно (\ref{phiN}).
\subsection{Блочная рекурсивная схема редукции размерности}
Для изложенной выше рекурсивной схемы предложено обобщение (блочная рекурсивная схема), которое комбинирует использование разверток и рекурсивной схемы с целью эффективного распараллеливания вычислений.
\par
Рассмотрим вектор \(y\) как вектор блочных переменных
\begin{displaymath}
y=(y_1,\dots,y_N)=(u_1,u_2,\dots,u_M)
\end{displaymath}
где \(i\)-я блочная переменная \(u_i\) представляет собой вектор размерности \(N\) из последовательно взятых компонент вектора \(y\), т.е. \(u_1=(y_1,y_2,\dots,y_{N_1}),u_2=(y_{N_1+1},y_{N_1+2},\dots,y_{N_1+N_2}),\dots,u_M=(y_{N-N_M+1},y_{N-N_M+2},\dots,y_N)\), причем \(N_1+N_2+\dots+N_M=N\).
\par
С использованием новых переменных основное соотношение многошаговой схемы (\ref{nestedScheme}) может быть переписано в виде
\begin{equation}
\label{blockNested}
\min\phi(y)_{y\in D}=\min_{u_1\in D_1}\min_{u_2\in D_2}\dots\min_{u_M\in D_M}\phi(y)
\end{equation}
где подобласти \(D_i,1\leqslant i\leqslant M\), являются проекциями исходной области поиска \(D\) на подпространства, соответствующие переменным \(u_1,1\leqslant i\leqslant M\).
\par
Формулы, определяющие способ решения задачи на основе соотношений (\ref{blockNested}) в целом совпадают с рекурсивной схемой (\ref{phiN})-(\ref{phiFirst}). Требуется лишь заменить исходные переменные \(y_i,1\leqslant i\leqslant N\), на блочные переменные \(u_1,1\leqslant i\leqslant M\).
\par
При этом принципиальным отличием от исходной схемы является тот факт, что в блочной схеме вложенные подзадачи
\begin{equation}
\label{subTasks}
\phi_i(u_1,\dots,u_i)=\min_{u_{i+1}\in D_{i+1}}\phi_{i+1}(u_1,\dots,u_i,u_{i+1}),1\leqslant i\leqslant M-1
\end{equation}
являются многомерными, и для их решения может быть применен способ редукции размерности на основе кривых Пеано.
\par
Число векторов и количество компонент в каждом векторе являются параметрами блочной многошаговой схемы и могут быть использованы для формирования подзадач с нужными свойствами. Например, если \(M=N\), т.е. \(y_i=u_i,1\leqslant i\leqslant N\), то блочная схема идентична исходной; каждая из вложенных подзадач является одномерной. А если \(M=1\), т.е. \(u=u_1=y\), то решение задачи эквивалентно ее решению с использованием единственной развертки, отображающей \([0,1]\) в \(D\); вложенные подзадачи отсутствуют.

\section{Реализация на Xeon Phi}
В 2012 году компания Intel представила первый сопроцессор с архитектурой Intel MIC (Intel® Many Integrated Core Architecture). Архитектура MIC позволяет использовать большое количество вычислительных ядер архитектуры x86 в одном процессоре. В результате для параллельного программирования могут быть использованы стандартные технологии, такие как OpenMP и MPI. 
\par
Архитектура Intel Xeon Phi поддерживает несколько режимов использования сопроцессора, которые можно комбинировать для достижения максимальной производительности в зависимости от характеристик решаемой задачи.
\par
В режиме Offload процессы MPI выполняются только на CPU, а на сопроцессоре происходит запуск отдельных функций, аналогично использованию графических ускорителей.
\par
В режиме MPI базовая система и каждый сопроцессор Intel Xeon Phi рассматриваются как отдельные равноправные узлы, и процессы MPI могут выполняться на центральных процессорах и сопроцессорах Xeon Phi в произвольных сочетаниях.
\subsection{Режим Offload}
Вначале рассмотрим ситуацию, когда проведение одного испытания является трудоемкой операцией. В этом случае ускоритель Xeon Phi может быть использован в режиме Offload для параллельного проведения сразу многих испытаний на одной итерации метода. Пересылки данных от CPU к Xeon Phi будут минимальны --- требуется лишь передать на сопроцессор координаты точек испытаний, и получить обратно значения функции в этих точках. Функции, определяющие обработку результатов испытаний в соответствии с алгоритмом и требующие работы с большим объемом накопленной поисковой информацией, могут быть эффективно реализованы на CPU. Общая схема организации вычислений с использованием Xeon Phi будет следующий.
\par
На CPU выполняются шаги 1 – 4 параллельного алгоритма глобального поиска из п. 2. При этом на каждой итерации происходит накопление координат точек испытания в буфере, и этот буфер передается на сопроцессор. На Xeon Phi выполняется параллельное вычисление значений функции в этих точках (шаг 5 алгоритма). Используется OpenMP распараллеливание цикла, на каждой итерации которого происходит вычисление значений функции. По завершению испытаний происходит передача вычисленных значений функции на CPU. 
\subsection{Режим MPI}
В случае, если проведение одного поискового испытания является относительно простой операцией, параллельное проведение многих итераций в режиме Offload не дает большого ускорения (сказывается влияние накладных расходов на передачу данных). Однако здесь можно увеличить вычислительную нагрузку на Xeon Phi, если применить блочную схему редукции размерности из п. 3.2, а для решения возникающих подзадач использовать сопроцессор в режиме MPI.
\par
Для организации параллельных вычислений будем использовать небольшое (2-3) число уровней вложенности в блочной схеме, при котором исходная задача большой размерности разбивается на 2-3 вложенные подзадачи меньшей размерности. Тогда, применяя в блочной рекурсивной схеме (\ref{blockNested}) для решения вложенных подзадач (\ref{subTasks}) параллельный алгоритм глобальной оптимизации, мы получим схему параллельных вычислений с широкой степенью вариативности (например, можно варьировать количество процессоров на различных уровнях оптимизации, т.е. при решении подзадач по различным переменным \(u_i\)).
\par
Общая схема организации вычислений с использованием нескольких узлов кластера и нескольких сопроцессоров состоит в следующем. Процессы параллельной программы образуют дерево, соответствующее уровням вложенных подзадач, при этом вложенные подзадачи
\begin{displaymath}
\phi_i(u_1,\dots,u_i)=\min_{u_{i+1}\in D_{i+1}}\phi_{i+1}(u_1,\dots,u_i,u_{i+1})
\end{displaymath}
при \(i=1,\dots,M-2\) решаются только с использованием CPU. Непосредственно в данных подзадачах вычислений значений оптимизируемой функции не происходит: вычисление значения функции \(\phi_i(u_1,\dots,u_i)\) --- это решение задачи минимизации следующего уровня. Каждая подзадача решается в отдельном процессе; обмен данными осуществляется лишь между процессами-предками и процессами-потомками.
\par
Подзадача последнего \((M–1)\)-го уровня 
\begin{displaymath}
\phi_{M-1}(u_1,\dots,u_{M-1})=\min_{u_{M}\in D_{M}}\phi_{M}(u_1,\dots,u_i,u_{M})
\end{displaymath} 
отличается от всех предыдущих подзадач --- в ней происходит вычисление значений оптимизируемой функции, т.к. \(\phi_M(u_1,\dots,u_M)=\phi(y_1,\dots,y_N)\). Подзадачи этого уровня решаются на сопроцессоре, и каждое ядро сопроцессора будет решать свою подзадачу \((M–1)\)-го уровня в отдельном MPI-процессе.
\par
Самый простой вариант использования данной схемы будет соответствовать двухкомпонентному вектору распараллеливания \(\pi=(\pi_1,\pi_2)\). Здесь \(\pi_1+1\) будет соответствовать числу MPI-процессов на CPU, а \(\pi_2\) --- числу MPI-процессов на Xeon Phi; тем самым общее количество процессов будет определяться как \(1+\pi_1+\pi_1\pi_2\).
\par
Отметим, что синхронный параллельный алгоритм глобальной оптимизации, описанный в п.3, в сочетании с блочной схемой редукции размерности обладает существенным недостатком, связанным с возможными простоями всех процессов, кроме корневого. Простои могут возникать в том случае, если часть потомков некоторого процесса закончили решение своих подзадач и отправили данные родителю раньше остальных. Для преодоления данной проблемы был применен предложенный в \cite{examinArtcle} асинхронный вариант параллельного алгоритма. Конкретные детали реализация блочной схемы в сочетании с асинхронным алгоритмом описаны в \cite{examinArtcle}.

\section{Результаты численных экспериментов}
Вычислительные эксперименты проводились на узле суперкомпьютера «Лобачевский», установленного в ННГУ им. Н.И. Лобачевского, на сегменте под управлением операционной системы семейства Windows (HPC Server 2008). Узел располагает двумя процессорами Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM, и двумя сопроцессорами Intel Xeon Phi 5110P.
\par
В работе \cite{gklsPaper} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п. Тестовые задачи, порождаемые данным генератором, характеризуются малым временем вычисления значений целевой функции. Поэтому с целью имитации вычислительной трудоемкости, присущей прикладным задачам оптимизации, расчет критерия был усложнен дополнительными вычислениями, не меняющими вид функции и расположение ее минимумов.
\begin{figure}[h!]
    \centering
		\includegraphics[width=0.45\textwidth]{isolines.png}
		\caption{Решение двумерной задачи}\label{fig:isolines}
\end{figure}
\par
Для примера на рис. \ref{fig:isolines} приведены линии уровня двумерной функции, порождаемой GKLS-генератором. Темные точки соответствуют 464 испытаниям, потребовавшимся для решения данной задачи с точностью 0.01 по координате; при этом число итераций составило 58 (на каждой итерации проводилось 8 испытаний). Линии уровня наглядно демонстрируют многоэкстремальность задачи, а расположение точек --- неравномерное покрытие, сгущающееся только в районе глобального минимума.
\par
Ниже приведены результаты численного сравнения трех последовательных алгоритмов --- DIRECT \cite{jonesLipOpt}, DIRECT\(l\)  \cite{gablonskyDirect} и алгоритма глобального поиска (АГП) (результаты работы первых двух алгоритмов приводятся по работе \cite{sergKvaDiaPaper}). Численное сравнение проводилось на классах функций Simple и Hard размерности 4 и 5 из \cite{sergKvaDiaPaper}. Глобальный минимум \(y^*\) считался найденным, если алгоритм генерировал точку испытания \(y_k\) в \(\delta\)-окрестности глобального минимума, т.е. \(\Vert y_k-y^*\Vert\leqslant\delta\). При этом размер окрестности выбирался (в соответствии с \cite{sergKvaDiaPaper}) как \(\delta=\Vert b-a\Vert \sqrt[N]{\Delta}\), здесь \(N\) --- размерность решаемой задачи, \(a\) и \(b\) --- границы области поиска \(D\), параметр \(\Delta=10^{-6}\) при \(N=4\) и \(\Delta=10^{-7}\) при \(N=5\). При использовании метода АГП для класса Simple выбирался параметр \(r=4.5\), для класса Hard --- \(r=5.6\); параметр построения кривой Пеано был фиксированный \(m=10\). Максимально допустимое число итераций составляло \(K_{max} = 1 000 000\).
\par
В таблице 1 отражено среднее число итераций \(k_{av}\), которые выполнил метод при решении серии задач из данных классов. Символ «>» отражает ситуацию, когда не все задачи класса были решены каким-либо методом. Это означает, что алгоритм был остановлен по причине достижения максимально допустимого числа итераций \(K_{max}\). В этом случае значение \(K_{max}=1 000 000\) использовалось при вычислении среднего значения числа итераций \(k_{av}\), что соответствует нижней оценке этого среднего значения. Количество нерешенных задач указано в скобках.
\begin{table}
    \centering
    \begin{tabular}{|c|p{3cm}|c|c|c|}
    \hline
    \(N\) & Класс функций & DIRECT & DIRECT\(l\) & АГП \\ \hline
    4 & Simple & >47282 (4) & 18983 & 11953\\ \cline{2-5}
      & Hard & >95708 (7) & 68754 & 68754\\ \hline
    5 & Simple & >16057 (1) & 16758 & 15920\\ \cline{2-5}
      & Hard & >217215 (16) & >269064 (4) & >148342 (4)\\ \hline
    \end{tabular}
    \caption{Среднее число итераций}
    \label{table:average_iters}
\end{table}
\par
Как видно из таблицы \ref{table:average_iters}, АГП превосходит методы DIRECT и DIRECT\(l\) на всех классах задач по среднему числу итераций. При этом в классе 5-Hard каждый из методов решил не все задачи: DIRECT не решил 16 задач, DIRECT\(l\) и АГП --- по 4 задачи.
\subsection{Задачи с трудоемким критерием}
Вначале приведем результаты экспериментов, которые соответствуют задачам с большим временем проведения одного поискового испытания. Оценим ускорение параллельного алгоритм глобального поиска (ПАГП), реализованного на CPU, в зависимости от использованного количества ядер \(p\). В таблицах \ref{table:time_speedUp_cpu} и \ref{table:iterations_speedUp_cpu} приведено ускорение по времени \(S(p)\) и по итерациям \(s(p)\). Ускорение вычислено относительно однопоточного запуска.
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
    & Simple & Hard & Simple & Hard \\ \hline
    2 & 2.45 & 2.20 & 1.15 & 1.32\\ \hline
	4 & 4.66 & 3.90 & 2.82 & 2.59\\ \hline
	8 & 7.13 & 7.35 & 3.47 & 5.34\\ \hline
    \end{tabular}
    \caption{Ускорение по времени \(S(p)\) на CPU}
    \label{table:time_speedUp_cpu}
\end{table}
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
    & Simple & Hard & Simple & Hard \\ \hline
	2 & 2.51 & 2.26 & 1.19 & 1.36\\ \hline
	4 & 5.04 & 4.23 & 3.06 & 2.86\\ \hline
	8 & 8.58 & 8.79 & 4.22 & 6.56\\ \hline
    \end{tabular}
    \caption{Ускорение по итерациям \(s(p)\) на CPU}
    \label{table:iterations_speedUp_cpu}
\end{table}
\par
Результаты экспериментов показывают значительное ускорение ПАГП при использовании CPU. Лучшие результаты получены при использовании всех ядер процессора.
\par
Далее изложим результаты экспериментов, проведенных с использованием Intel Xeon Phi. Вначале рассмотрим эксперименты на одном Xeon Phi в режиме Offload. Варьировалось количество потоков на сопроцессоре, все остальные параметры совпадают с предыдущими запусками. Приведено (таблицы \ref{table:time_speedUp_phi} и \ref{table:iterations_speedUp_phi}) ускорение относительно восьмипоточного запуска на центральном процессоре.
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
    & Simple & Hard & Simple & Hard \\ \hline
	60 & 0.54 & 1.02 & 1.07 & 1.61\\ \hline
	120 & 0.55 & 1.17 & 1.05 & 2.61\\ \hline
	240 & 0.51 & 1.06 & 1.07 & 4.17\\ \hline
    \end{tabular}
    \caption{Ускорение по времени \(S(p)\) на Phi}
    \label{table:time_speedUp_phi}
\end{table}
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
    & Simple & Hard & Simple & Hard \\ \hline
	60 & 8.13 & 7.32 & 9.87 & 6.55 \\ \hline
	120 & 16.33 & 15.82 & 15.15 & 17.31 \\ \hline
	240 & 33.07 & 27.79 & 38.80 & 59.31 \\ \hline
    \end{tabular}
    \caption{Ускорение по итерациям \(s(p)\) на Phi}
    \label{table:iterations_speedUp_phi}
\end{table}
\par
Результаты экспериментов показывают, что только на классе Simple при \(N=4\) реализация на Xeon Phi медленнее, чем CPU реализация, на классе Hard при \(N=4\) и на классе Simple при \(N=5\) версии для Xeon Phi и CPU демонстрирует примерно равную эффективность, а на классе Hard при \(N=5\) версия для Xeon Phi существенно превосходит CPU-реализацию. Наибольшее преимущество реализация для Xeon Phi показывает при 120 потоках для четырехмерных задач и при 240 потоках для пятимерных. При этом на всех классах наблюдается значительное ускорение по итерациям, а также практически линейная масштабируемость от числа потоков.
\par
Далее рассмотрим эксперименты с запуском сопроцессора в режиме MPI. Число уровней вложенности разбиваемой задачи равно двум. Число процессов, запущенных на сопроцессоре, –-- 30 (таблицы \ref{table:time_speedUp_phi_30mpi} и \ref{table:iterations_speedUp_phi_30mpi}) и 60 (таблицы \ref{table:time_speedUp_phi_60mpi} и \ref{table:iterations_speedUp_phi_60mpi}), что соответствует ситуации, когда у корня дерева имеется соответственно 30 и 60 потомков. Процессы, работающие на сопроцессоре, использовали OpenMP для параллельного вычисления значений функции. Ускорение приведено относительно восьмипоточного запуска на CPU.
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
    & Simple & Hard & Simple & Hard \\ \hline
	4 & 3,18 & 5,14 & 5,15 & 14,65 \\ \hline
	8 & 3,53 & 1,49 & 3,35 & 10,49 \\ \hline
	16 & 1,09 & 2,77 & 4,61 & 14,36 \\ \hline
    \end{tabular}
    \caption{Ускорение по времени \(S(p)\) на Phi, 30 MPI процессов}
    \label{table:time_speedUp_phi_30mpi}
\end{table}
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
	4 & 11,19 & 19,19 & 9,27 & 27,38 \\ \hline
	8 & 13,97 & 3,63 & 8,24 & 23,85 \\ \hline
	16 & 1,29 & 3,56 & 8,28 & 22,12 \\ \hline
	\end{tabular}
    \caption{Ускорение по итерациям \(s(p)\) на Phi, 30 MPI процессов}
    \label{table:iterations_speedUp_phi_30mpi}
\end{table}
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
    & Simple & Hard & Simple & Hard \\ \hline
	4 & 2,76 & 3,20 & 4,19 & 16,56 \\ \hline
	8 & 3,05 & 7,73 & 4,08 & 14,87 \\ \hline
	16 & 0,89 & 5,21 & 2,72 & 13,05 \\ \hline
    \end{tabular}
    \caption{Ускорение по времени \(S(p)\) на Phi, 60 MPI процессов}
    \label{table:time_speedUp_phi_60mpi}
\end{table}
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
	4 & 10,36 & 3,34 & 6,01 & 26,04 \\ \hline
	8 & 9,04 & 28,81 & 7,68 & 25,02 \\ \hline
	16 & 1,27 & 28,66 & 7,19 & 30,38 \\ \hline
	\end{tabular}
    \caption{Ускорение по итерациям \(s(p)\) на Phi, 60 MPI процессов}
    \label{table:iterations_speedUp_phi_60mpi}
\end{table}
\par
Результаты экспериментов показывают, что запуск в режиме MPI значительно превосходит Offload режим по времени работы, однако проигрывает по числу итераций. Лучшие результаты для простого класса задач наблюдаются при использовании 30 процессов на сопроцессоре, а для сложного класса, при 60 процессах. Таким образом, можно сделать вывод, что задачи с большим временем работы, для которых важно число итераций метода оптимизации, лучше использовать запуск в режиме Offload, а для более быстро решаемых задач, лучше подходит режим MPI. 
\subsection{Задачи с легко вычисляемым критерием}
Рассмотрим теперь решение задач, в которых поисковые испытания проводятся быстро. Вначале приведем результаты OpenMP версии. В таблице \ref{table:time_speedUp_cpu_2} приведено ускорение относительно однопоточного запуска.
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
    & Simple & Hard & Simple & Hard \\ \hline
	2 & 1,09 & 1,09 & 0,56 & 0,57 \\ \hline
	4 & 1,10 & 1,06 & 0,74 & 0,66 \\ \hline
	8 & 0,90 & 1,11 & 0,45 & 1,77 \\ \hline
	16 & 0,66 & 0,92 & 0,76 & 0,66 \\ \hline
	32 & 0,29 & 0,38 & 0,22 & 0,55 \\ \hline
    \end{tabular}
    \caption{Ускорение по времени \(S(p)\) на CPU}
    \label{table:time_speedUp_cpu_2}
\end{table}
\par
Эксперименты показывают незначительное ускорение, а во многих случаях и замедление при вычислениях только на центральном процессоре.
\par
Далее рассмотрим ускорение при использовании Xeon Phi в режиме MPI. Результаты приведены относительно однопоточного запуска на CPU. В таблице \ref{table:time_speedUp_phi_30mpi_easy} приведено ускорение по времени \(S(p)\) для запуска 30 процессов на Xeon Phi, в таблице \ref{table:time_speedUp_phi_60mpi_easy} для 60 процессов.
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
    & Simple & Hard & Simple & Hard \\ \hline
	1 & 0,35 & 2,28 & 0,37 & 1,92 \\ \hline
	2 & 0,94 & 1,95 & 0,39 & 1,23 \\ \hline
	4 & 1,23 & 0,43 & 0,54 & 2,63 \\ \hline
	8 & 0,85 & 0,56 & 0,40 & 1,53 \\ \hline
	\end{tabular}
    \caption{Ускорение по времени \(S(p)\) на Phi, 30 MPI процессов}
    \label{table:time_speedUp_phi_30mpi_easy}
\end{table}
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \(p\) & \multicolumn{2}{|c|}{\(N=4\)} & \multicolumn{2}{|c|}{\(N=5\)}\\ \cline{2-5}
    & Simple & Hard & Simple & Hard \\ \hline
	1 & 0,59 & 2,56 & 1,22 & 3,30 \\ \hline
	2 & 0,89 & 0,87 & 1,29 & 2,76 \\ \hline
	4 & 1,09 & 1,03 & 0,38 & 5,32 \\ \hline
	8 & 0,70 & 0,78 & 0,69 & 2,19 \\ \hline
	\end{tabular}
    \caption{Ускорение по времени \(S(p)\) на Phi, 60 MPI процессов}
    \label{table:time_speedUp_phi_60mpi_easy}
\end{table}
\par
Результаты экспериментов показывают ускорение на Xeon Phi, превосходящее ускорение на центральном процессоре. Лучшие результаты получены при использовании 30 процессов на Xeon Phi только для простых четырехмерных задач, все остальные задачи лучше всего решаются при использовании 60 процессов.
\par
Теперь перейдем к эксперименту с шестимерными задачами простого класса. При решении шестимерной задачи с использованием одного процесса (вложенные подзадачи отсутствуют) и сохранением точности построения развертки \(m=10\) требуется использовать числа с плавающей запятой расширенной точности, что ведет к значительному увеличению времени работы алгоритма. При этом в случае использования хотя бы одного уровня вложенных подзадач, задачи на каждом уровне будут решаться без подключения расширенной точности, что обеспечивает дополнительное преимущество блочной многошаговой схемы.
\par
Рассмотрим результаты эксперимента с использованием Xeon Phi в режиме MPI, таблица \ref{table:time_speedUp_phi_2}. Число уровней вложенности разбиваемой задачи равно двум. Приведены результаты экспериментов для разбиений 3:3 (три переменных на первом уровне, три –-- на втором) и 4:2 (4 переменных на первом уровне, две –-- на втором). Число процессов, запущенных на сопроцессоре равно 30 и 60, что соответствует тому, что у корня дерева имеется соответственно 30 и 60 потомков. Процессы, работающие на Xeon Phi, использовали OpenMP для параллельного вычисления значений функции. Ускорение приведено относительно однопоточного запуска на CPU.
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \(p\) & \multicolumn{2}{|c|}{30 процессов} & \multicolumn{2}{|c|}{60 процессов}\\ \cline{2-5}
	& 3:3 & 4:2 & 3:3 & 4:2 \\ \hline
	2 & 7,30 & 9,87 & 17,68 & 18,21 \\ \hline
	4 & 5,99 & 8,18 & 6,60 & 13,57 \\ \hline
	8 & 3,14 & 12,74 & 5,92 & 25,58 \\ \hline
	16 & 3,54 & 7,61 & 4,87 & 11,33 \\ \hline
	32 & 3,41 & 6,85 & 3,73 & 6,50 \\ \hline
\end{tabular}
    \caption{Ускорение по времени \(S(p)\) на Phi}
    \label{table:time_speedUp_phi_2}
\end{table}
\par
Результаты экспериментов показывают значительное ускорение запуска на сопроцессоре по сравнению с запуском только на центральном процессоре. Лучшее ускорение наблюдается при шестидесяти MPI процессах на Xeon Phi по 8 потоков на каждый процесс, время работы в 25 раз меньше, чем решение задачи только на центральном процессоре.

\section{Заключение}
В работе рассмотрен параллельный алгоритм глобального поиска, разработанный в рамках информационно статистического подхода. Предложен подход к распараллеливанию данного алгоритма, который может быть эффективен как в задачах с трудоемким критерием оптимизации, так и в задачах с относительно простым критерием. Проведены эксперименты с реализацией предложенного параллельного алгоритма на суперкомпьютере «Лобачевский» с использованием сопроцессоров Intel Xeon Phi. Результаты вычислительных экспериментов подтверждают высокую эффективность распараллеливания для Xeon Phi во всех классах рассмотренных задач.

\begin{thebibliography}{10}

\bibitem{evtushenko}
Евтушенко Ю.Г.
\newblock {\em Методы решения экстремальных
  задач и их применение в системах
  оптимизации}.
\newblock М.: Наука, 1982, 432с.

\bibitem{pinter}
Pinter J.
\newblock {\em Global optimization in action (Continuous and Lipschitz
  optimization: algorithms, imple-mentations and applications)}.
\newblock Kluwer Academic Publishers, 1996, 507 p.

\bibitem{sergKvaDiag}
Квасов~Д.Е. Сергеев~Я.Д.
\newblock {\em Диагональные методы глобальной
  оптимизации}.
\newblock М.: Физматлит, 2008, 352c.

\bibitem{strSergOptBook}
Sergeyev~Ya.D. Strongin~R.G.
\newblock {\em Global Optimization with Non-convex Constraints. Sequential and
  Parallel Algorithms}.
\newblock Kluwer Academic Publishers, 2000, 704p.

\bibitem{strGergrParOptBook}
Стронгин Р.Г. Гергель В.П. Гришагин
  В.А.~Баркалов К.А.
\newblock {\em Параллельные вычисления в задачах
  глобальной оптимизации}.
\newblock М.: Издательство Московского
  университета, 2013, 280с.

\bibitem{sergKvaDiaPaper}
Sergeyev Ya.D.~Kvasov D.E.
\newblock Global search based on efficient diagonal partitions and a set of
  lipschitz constants.
\newblock {\em SIAM Journal on Optimization}, 16(3):910–937, 2006.

\bibitem{zilinsk}
Žilinskas J.
\newblock Branch and bound with simplicial partitions for global optimization.
\newblock {\em Mathematical Modelling and Analysis}, 13(1):145–159, 2008.

\bibitem{gorodGrishOptBook}
Гришагин~В.А Городецкий~С.Ю.
\newblock {\em Нелинейное программирование и
  многоэкстремальная оптимизация}.
\newblock Н. Новгород: Изд-во ННГУ, 2007.

\bibitem{examinArtcle}
Сысоев А.В. Баркалов К.А. Гергель
  В.П.~Лебедев И.Г.
\newblock Mpi-реализация блочной многошаговой
  схемы параллельного решения задач
  глобальной оптимизации.
\newblock In {\em Суперкомпьютерные дни в России:
  Труды международной конференции (28-29
  сентября 2015 г., г. Москва)}, pages 411--419. М.:
  Изд-во МГУ, 2015.

\bibitem{gklsPaper}
Gaviano M. Lera D. Kvasov D.E.~Sergeyev Ya.D.
\newblock Software for generation of classes of test functions with known local
  and global minima for global optimization.
\newblock {\em ACM Transactions on Mathematical Software}, 29:469--480, 2003.

\bibitem{jonesLipOpt}
Jones D.R. Perttunen C.D.~Stuckman B.E.
\newblock Lipschitzian optimization without the lipschitz constant.
\newblock {\em Journal of Optimization Theory and Applications},
  79(1):157–181, 1993.

\bibitem{gablonskyDirect}
Kelley~C.T. Gablonsky~J.M.
\newblock A locally-biased form of the direct algorithm.
\newblock {\em Journal of Global Optimization}, 21(1):27--37, 2001.

\end{thebibliography}

\end{document}

### 2016-03-04 2016/PAVT/numerical_experiments.tex
Move bibliography to main_single.tex main_single.tex and main_eng.tex are ready for sending to journal

@@
 Как видно из таблицы \ref{table:average_iters}, АГП превосходит методы DIRECT и DIRECT\(l\) на всех классах задач по среднему числу итераций.
-При этом в классе 5-Hard каждый из методов решил не все задачи: DIRECT не решил 16 задач, DIRECT\(l\) и АГП – по 4 задачи.
+При этом в классе 5-Hard каждый из методов решил не все задачи: DIRECT не решил 16 задач, DIRECT\(l\) и АГП --- по 4 задачи.
 \subsection{Задачи с трудоемким критерием}
@@
 Число уровней вложенности разбиваемой задачи равно двум.
-Число процессов, запущенных на сопроцессоре, –-- 30 (таблицы \ref{table:time_speedUp_phi_30mpi} и \ref{table:iterations_speedUp_phi_30mpi}) и 60(таблицы \ref{table:time_speedUp_phi_60mpi} и \ref{table:iterations_speedUp_phi_60mpi}), что соответствует ситуации, когда у корня дерева имеется соответственно 30 и 60 потомков.
+Число процессов, запущенных на сопроцессоре, –-- 30 (таблицы \ref{table:time_speedUp_phi_30mpi} и \ref{table:iterations_speedUp_phi_30mpi}) и 60 (таблицы \ref{table:time_speedUp_phi_60mpi} и \ref{table:iterations_speedUp_phi_60mpi}), что соответствует ситуации, когда у корня дерева имеется соответственно 30 и 60 потомков.
 Процессы, работающие на сопроцессоре, использовали OpenMP для параллельного вычисления значений функции.
@@
 \par
-Эксперименты показывают незначительное ускорение, а во многих случаях – и замедление при вычислениях только на центральном процессоре.
+Эксперименты показывают незначительное ускорение, а во многих случаях и замедление при вычислениях только на центральном процессоре.
 \par
@@
 Число уровней вложенности разбиваемой задачи равно двум.
-Приведены результаты экспериментов для разбиений 3:3 (три переменных на первом уровне, три – на втором) и 4:2 (4 переменных на первом уровне, две – на втором).
+Приведены результаты экспериментов для разбиений 3:3 (три переменных на первом уровне, три –-- на втором) и 4:2 (4 переменных на первом уровне, две –-- на втором).
 Число процессов, запущенных на сопроцессоре равно 30 и 60, что соответствует тому, что у корня дерева имеется соответственно 30 и 60 потомков.

### 2016-03-05 2016/PAVT/main.tex
Finish introduction.tex
Finish chapter parallel_alg.tex
Update of numerical_experiments.tex
Add english abstract and refs
Fixes in PAVT paper

@@
 \usepackage{indentfirst}
-\usepackage{color}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
+\usepackage{multirow}
+\usepackage{fancybox,fancyhdr}
 
@@
 
-\usepackage[backend=biber]{biblatex}
-\addbibresource{refs.bib}
+\bibliographystyle{unsrt}
 
-\newcommand{\tbd}[1]{
-    \textcolor{red}{#1}
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator{\sign}{sign}
+\fontsize{10}{12pt}\selectfont
+
+\fancypagestyle{firststyle}
+{
+\fancyhead[L]{УДК 519.853.4}
+\renewcommand{\headrulewidth}{0pt}
 }
-\newcommand{\tworowcell}[2]{
-    \begin{tabular}{@{}c@{}}#1 \\ #2\end{tabular}
-}
-\renewcommand{\setminus}{\mathbin{\backslash}}
 
@@
 \footnote{
-Исследование выполнено при финансовой поддержке РФФИ в рамках научного проекта № 16-31-00244 мол\_а «Параллельные методы решения вычислительно трудоемких задач глобальной оптимизации на гибридных кластерных системах»}}
+Исследование выполнено при финансовой поддержке РФФИ в рамках научного проекта № 16-31-00244 мол\_а «Параллельные методы решения вычислительно трудоемких задач глобальной оптимизации на гибридных кластерных системах»}
+\footnote{
+Статья вошла в программу конференции "Параллельные вычислительные технологии (ПаВТ) 2016"}}
 \author{К.А. Баркалов, И.Г. Лебедев, В.В. Соврасов, А.В. Сысоев}
@@
 \maketitle
-
+\thispagestyle{firststyle}
 \begin{abstract}
@@
 \input{conclusion}
-\printbibliography
+
+\bibliography{refs}
+
 \end{document}

### 2016-03-05 2016/PAVT/main_eng.tex
Fixes in PAVT paper

@@
 \usepackage{indentfirst}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
+\usepackage{fancybox,fancyhdr}
 
@@
 
+\fancypagestyle{firststyle}
+{
+\fancyhead[L]{UDC 519.853.4}
+\renewcommand{\headrulewidth}{0pt}
+}
+
 \begin{document}
@@
 \footnote{
-The paper was supported by the Russian Science Foundation, project No 15-11-30022 “Global optimization, supercomputing  computations, and applications”}}
+The paper was supported by the Russian Science Foundation, project No 15-11-30022 “Global optimization, supercomputing  computations, and applications”}
+\footnote{
+The paper was presented at the conference "Parallel computational technologies (PCT) 2016"}}
 \author{K.A. Barkalov, I.G. Lebedev, V.V. Sovrasov, A.V. Sysoyev}

### 2016-03-05 2016/PAVT/main_single.tex
Fixes in PAVT paper

@@
 \usepackage{multirow}
-\usepackage{graphicx}
+\usepackage{fancybox,fancyhdr}
 
@@
 \fontsize{10}{12pt}\selectfont
+
+\fancypagestyle{firststyle}
+{
+\fancyhead[L]{УДК 519.853.4}
+\renewcommand{\headrulewidth}{0pt}
+}
+
 \begin{document}
-
 \title{Реализация параллельного алгоритма поиска глобального экстремума функции на Intel Xeon Phi\
 \footnote{
-Исследование выполнено при финансовой поддержке РФФИ в рамках научного проекта № 16-31-00244 мол\_а «Параллельные методы решения вычислительно трудоемких задач глобальной оптимизации на гибридных кластерных системах»}}
+Исследование выполнено при финансовой поддержке РФФИ в рамках научного проекта № 16-31-00244 мол\_а «Параллельные методы решения вычислительно трудоемких задач глобальной оптимизации на гибридных кластерных системах»}
+\footnote{
+Статья вошла в программу конференции "Параллельные вычислительные технологии (ПаВТ) 2016"}}
 \author{К.А. Баркалов, И.Г. Лебедев, В.В. Соврасов, А.В. Сысоев}
@@
     \centering
-		\includegraphics[width=0.45\textwidth]{isolines.png}
+		\includegraphics[width=0.45\textwidth]{isolines.eps}
 		\caption{Решение двумерной задачи}\label{fig:isolines}

### 2016-05-02 2016/NUMTA/paper.tex
Add tex version of NUMTA paper

\documentclass{aip-cp}

\usepackage[numbers]{natbib}
\usepackage{rotating}
\usepackage{graphicx}

% Document starts
\begin{document}

% Title portion
\title{Local Tuning in Multilevel Scheme of Parallel Global Optimization}

\author[aff1]{Konstantin Barkalov\corref{cor1}}
\author[aff1]{Ilya Lebedev}

\affil[aff1]{University of Nizhni Novgorod, Nizhni Novgorod, Russia.}
\corresp[cor1]{Corresponding author: barkalov@vmk.unn.ru}

\maketitle

\begin{abstract}
This work considers a parallel algorithm for solving the multidimensional multiextremal optimization problems. The algorithm combines nested optimization scheme and Peano-type space filling curves for the dimensionality reduction. To decrease the number of iterations of the global algorithm, we use one of local tuning techniques based on the adaptive estimation of the global optimizer. Parallel algorithm with mixed local-global strategy of search is proposed as well. The efficiency of the parallel algorithm was investigated using a supercomputer. The speedup of the algorithm using several nodes as compared with the serial algorithm has been demonstrated experimentally.
\end{abstract}

% Head 1
\section{INTRODUCTION}

This paper considers the problem of multidimensional multiextremal optimization
\begin{displaymath}
\phi(y^*)=\min\{\phi(y):y\in D\}
\end{displaymath}
\begin{displaymath}
D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}
\end{displaymath}
where $a,b \in R^N$ are given vectors. As for the class of the considered problems, the fulfillment of the two following important conditions is presupposed.

First, it is supposed that the objective function $\phi(y)$ can be defined non-analytically, but using an algorithm for calculation of its values at some points of domain $D$. In this case, such a calculation (\textit{trial}) is computationally intensive.

Second, we shall assume that $\phi(y)$ satisfies Lipschitz condition
\begin{displaymath}
|\phi(y_1)-\phi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
\end{displaymath}
that is typical for the applied problems [1–4].

Solving the multiextremal optimization problems is significantly more computationally intensive as compared with other types of optimization problems, since the global optimum is an integral characteristic of the problem solved and requires an analysis of the whole search domain. The efficient algorithms for solving the problems of this kind have been developed in Lobachevsky State University of Nizhnii Novgorod within the framework of the information-statistical approach [5, 6]. At present, the parallelization of the algorithms is the main direction of further development of these ones since the use of the supercomputer systems and of the parallel computing technologies allows expanding the field of application of the global optimization algorithms essentially [7 – 12].

This paper continues the series of studies the initial results of which were provided in [13,14].

\section{PARALLEL ALGORITHM FOR SOLVING THE MULTIDIMENSIONAL PROBLEMS}

% Head 2
\subsection{DIMENSION REDUCTION}
There is a number of ways to adapt effective one-dimensional algorithms for solving multidimensional problems; see, for example, the diagonal partitions method in [15,16] or the simplicial partitions method in [17].

In the present study, we will use a mixed scheme of the dimension reduction utilizing Peano curve and the nested optimization scheme. Such schemes (used separately) allow reducing the solving of the multidimensional optimization problems to the solving of one or several linked one-dimensional optimization problems.

The use of Peano-type space filling curves (\textit{evolvents}) $y(x)$, which map a unit interval $[0,1]$ onto an $n$-dimensional cube continuously and unambiguously allows the reduction of the problem of minimization in domain $D$ to a minimization problem on the interval $[0,1]$
\begin{equation}
\label{evolvent}
\phi(y^*)=\phi(y(x^*))=\min\{\phi(y(x)):x\in [0,1]\}
\end{equation}
The second method for the dimension reduction is the multilevel scheme of nested optimization based on the relation
\begin{equation}
\label{nestedScheme}
\min_{y\in D}\phi(y) =\min_{u_1\in D_1}\min_{u_2\in D_2}\dots\min_{u_M\in D_M }\phi(y)
\end{equation}
where the $i^{th}$ variable ui is a vector of the dimensionality $N_i$ of the components of vector $y$ taken sequentially i. e.
\begin{equation}
y=(y_1, y_2, \dots ,y_N)=(u_1,u_2,\dots,u_M),
\end{equation}

In scheme (\ref{nestedScheme}), the nested subproblems are the multidimensional ones, and the dimension reduction method based on Peano curves can be applied for solving of these ones. The number of vectors $u_i$ and the quantity of components in each vector are the parameters of the block recursion scheme and can be utilized for the forming of the subproblems with necessary properties. For example, if $M=N$ i. e. $u_i=u_i,1\leqslant i\leqslant N$ , then each nested subproblem is a one-dimesional one. If $M=1$ i. e. $u=u_1=y$ , then solving the problem is an equivalent to its solving using a single evolvent mapping $[0,1]$ onto $D$; the nested subproblems are absent.
The issues of the numerical construction of the Peano curve approximations and of using the nested optimization scheme are considered in details in [3,4].

\subsection{PARALLEL GLOBAL SEARCH ALGORITHM}
To solve the one-dimensional subproblems in the nested optimization scheme, we will use characteristical global search algorithm (GSA).
The first iteration $x^1$ is executed in an arbitrary internal point of the search interval. The choice of the next iteration pointv$x^{k+1}$, $k\geqslant 1$ is performed according to the following rules.

\begin{enumerate}
\item To arrange the search information in the order of increasing values of the iteration points executed earlier i. e. $x_1\leqslant x_2 \leqslant \dots \leqslant x_k$.
\item To compute a quantity $R(i)$ called a \textit{characteristic} of the interval for each interval $(x_{i-1},x_i),1\leqslant i \leqslant k$.
\item To determine the interval $(x_{t-1},x_t)$ which the maximum characteristic $R(t)=\max{\{R(i): 1\leqslant i \leqslant k}\}$ corresponds to.
\item To compute the next iteration point $x^{k+1}$ in the interval with the maximum characteristic i. e. $x^{k+1}\in (x_{t-1},x_t)$.
\end{enumerate}

A detailed description of various variants of this algorithm is presented in [5, 18, 19, 20].

One of the methods of the parallelization of the characteristic algorithms is based on the following consideration. The characteristic of an interval $R(t)$ can be considered as a measure of importance of this interval for the following search for the solution in this one. Therefore, after the selection of the trial point for the first processor in full accordance with the sequential algorithm. It is reasonable to choose the trial point for the second processor from the next interval in importance (i. e. from the interval with the next value of characteristic in the decreasing order), etc. A parallel algorithm constructed in such a way would differ from its sequential prototype by rules 3 and 4 only, namely:
\begin{enumerate}
\item[3.] To arrange the characteristics $R(i),1\leqslant i\leqslant \tau$  in the decreasing order, i. e. $R(t_1)\geqslant R(t_2)\geqslant \dots \geqslant R(t_{k-1}) \geqslant R(t_k)$
\item[4.] To select $p$ highest characteristics with the indices $t_j,1\leqslant j \leqslant p$ and to execute the trials in the points
\begin{displaymath}
x^{k+j} \in (x_{t_j - 1}, x_{t_j}),1\leqslant j \leqslant p
\end{displaymath}
\end{enumerate}

in the intervals corresponding to these characteristics.

Various variants of the parallel algorithms based on this idea are presented in [7-12].

% Head 3
\subsubsection{MIXED LOCAL-GLOBAL ALGORITHM}
A scheme aimed at the acceleration of the search process is outlined in this subsection. The idea of acceleration is to magnify the characteristics of the intervals containing the best current estimates by introducing some factor depending on the function values estimated at the end points of the corresponding interval. Following this idea, the “local” characteristics of the intervals can be introduced [5, 21].

Next, it is possible to build a scheme mixing the 'global' and the 'local' decision rules, i. e. switching between these ones in some systematic way. In our experiments, we use global-to-local ratio q, specifying the number of global trials preceding each local trial.

The mixed strategy has the following features:

\begin{itemize}
\item both decision rules are based on the same information, so that each decision action (no matter local or global) uses the results of all the trials performed.
\item non-stop global search assures the global convergence; the aim of the local refinement is to accelerate the attainment of low function values.
\end{itemize}

\section{NUMERICAL EXPERIMENTS}
Computing experiments were carried out on Lobachevsky supercomputer. Each supercomputer node included two Intel Sandy Bridge E5-2660 2.2 GHz CPU and 64 Gb RAM. Each CPU had 8 cores (total 16 cores per a node). For the implementation of the parallel algorithm, Intel C++ Compiler 14.0.2 was used.

The problems generated by GKLS generator [22] were used as the test ones. This generator allows generating the multiextremal optimization problems with the properties known in advance: number of local minima, size of attraction domains, global minimum point, function values, etc. The experiment included solving a series of 200 problems with $N=4$ and $N=5$. The global minimum $y^*$ was considered to be found, if the algorithm generates a trial point $y^k$ in its $\delta$-vicinity $\Vert y^k-y^*\Vert\leqslant \delta$. The size of the vicinity was selected as $\delta=0.01\Vert b- a\Vert$ where $a$ and $b$ were the borders of the search domain $D$. The maximum allowable number of trials was $K_{max} = 10^6$.

First, let us consider the results of experiments corresponding to the sequential algorithm. Let us compare initial GSA and mixed local-global algorithm. The CPU time consumption for the solving of four series of problems from Simple and Hard classes are presented in Table \ref{tab:mixedResults}(a) and the numbers of the executed search trials subject to global-to-local ratio $q$ are presented in Table \ref{tab:mixedResults}(b). The row denoted as "---" corresponds to the algorithm without the local refining. The results demonstrate the reduction of the computation time as well as of the number of iterations when using the local-global search strategy. Also, it is seen that the best results have been achieved at the mixing parameters $q=2$ and $q=4$. The less and greater values of this parameter were insufficient. Therefore, the results of investigation of the mixed parallel algorithm will be presented for the value of parameter $q=4$.

\begin{table}[h]
	\caption{}
	\label{tab:mixedResults}
	\begin{minipage}{0.49\linewidth}
	\begin{tabular}{cccccc}
		\multicolumn{6}{c}{(a) CPU time consumption }\\
		\hline\noalign{\smallskip}
		$q$ & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & Simple & Hard & & Simple & Hard  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}		
--- &	0.08253 & 0.24744 & & 0.11036 &	1.85801 \\
 1  &	0.0875  & 0.18375 & & 0.08196 &	0.90628 \\
 2	&   0.07698	& 0.15921 & & 0.07491 & 0.82082 \\
 4  &   0.07515 & 0.14893 & & 0.07596 & 0.93878 \\
 8  &   0.07737 & 0.14959 & & 0.07537 & 1.06467 \\
16  &   0.07419 & 0.17232 & & 0.08508 & 1.51871 \\
		\noalign{\smallskip}\hline
	\end{tabular} 
	\end{minipage}

	\begin{minipage}{0.49\linewidth}
	\begin{tabular}{cccccc}
		\multicolumn{6}{c}{(b) Number of iterations}\\
		\hline\noalign{\smallskip}
		$q$ & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & Simple & Hard & & Simple & Hard  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
--- &  12558 & 37410 & & 15226 & 248275 \\
1   &  11370 & 25956 & &  9959 & 120978 \\
2	&   9842 & 22456 & &  9047 & 109995 \\
4   &   9574 & 20942 & &  9018 & 121934 \\
8   &   9936 & 20879 & &  8940 & 136174 \\
16	&   9692 & 23992 & & 10410 & 179490 \\
		\noalign{\smallskip}\hline
	\end{tabular}
	\end{minipage}
\end{table}

The computation times for the mixed parallel algorithm subject to the number of processors employed $p$ are presented in Table \ref{tab:parallelResults}(a), and the ыpeedup of the parallel algorithm relative to the sequential one is shown in Table \ref{tab:parallelResults}(b).

\begin{table}[h]
	\caption{}
	\label{tab:parallelResults}
	\begin{minipage}{0.49\linewidth}
	\begin{tabular}{cccccc}
		\multicolumn{6}{c}{(a) CPU time consumption }\\
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & Simple & Hard & & Simple & Hard  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
1    &  0.083 & 0.247 & & 0.110 & 1.858 \\
1+1  &  0.043 & 0.084 & & 0.159 & 0.575 \\
2+1  &  0.039 & 0.058 & & 0.091 & 0.316 \\
4+1	 &  0.058 & 0.039 & & 0.055 & 0.171 \\
8+1  &  0.031 & 0.036 & & 0.048 & 0.103 \\
16+1 &  0.061 & 0.061 & & 0.071 & 0.129 \\
		\noalign{\smallskip}\hline
	\end{tabular} 
	\end{minipage}

	\begin{minipage}{0.49\linewidth}
	\begin{tabular}{cccccc}
		\multicolumn{6}{c}{(b) Speedup on CPU}\\
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & Simple & Hard & & Simple & Hard  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
2  &  1.91 & 2.94 & & 0.70 &  3.23 \\
3  &  2.12 & 4.27 & & 1.22 &  5.88 \\
5  &  1.42 & 6.28 & & 2.01 & 10.89 \\
6  &  2.66 & 6.88 & & 2.28 & 18.05 \\
17 &  1.34 & 4.03 & & 1.54 & 14.40 \\
		\noalign{\smallskip}\hline
	\end{tabular}
	\end{minipage}
\end{table}

The results of the experiments have shown the mixed local-global algorithm combined with the multilevel scheme of dimension reduction to demonstrate a good speedup, the best speedup has been achieved when solving the complex problems of 5-Hard class. Therefore, the proposed algorithm can be applied for the parallel solving of the multidimensional global optimization problems using the state-of-art supercomputer systems.

\section{ACKNOWLEDGMENTS}
This research was supported by the Russian Science Foundation, project No 16-11-10150 “Novel efficient methods and software tools for the time consuming decision making problems with using supercomputers of superior performance”.

% References

\nocite{*}
\bibliographystyle{aipnum-cp}%
\bibliography{paper}%

\end{document}

### 2016-05-02 2016/NUMTA/paper.tex
Finish tex version of NUMTA paper

@@
 
-% Document starts
 \begin{document}
 
-% Title portion
 \title{Local Tuning in Multilevel Scheme of Parallel Global Optimization}
@@
 \end{displaymath}
-that is typical for the applied problems [1–4].
+that is typical for the applied problems \citep{automaticaAppl,svmAppl,recogAppl,determApproaches}.
 
 Solving the multiextremal optimization problems is significantly more computationally intensive as compared with other types of optimization problems, since the global optimum is an integral characteristic of the problem solved and requires an analysis of the whole search domain.
-The efficient algorithms for solving the problems of this kind have been developed in Lobachevsky State University of Nizhnii Novgorod within the framework of the information-statistical approach [5, 6].
-At present, the parallelization of the algorithms is the main direction of further development of these ones since the use of the supercomputer systems and of the parallel computing technologies allows expanding the field of application of the global optimization algorithms essentially [7 – 12].
+The efficient algorithms for solving the problems of this kind have been developed in Lobachevsky State University of Nizhnii Novgorod within the framework of the information-statistical approach \citep{strongSerg,strongSergLera}.
+At present, the parallelization of the algorithms is the main direction of further development of these ones since the use of the supercomputer systems and of the parallel computing technologies allows expanding the field of application of the global optimization algorithms essentially \citep{parAlg,parAlg2,gergelLNCS,gergelStrParComputing,mixedAlgLNCS,barkXeon}.
 
-This paper continues the series of studies the initial results of which were provided in [13,14].
+This paper continues the series of studies the initial results of which were provided in \citep{barkGergNestedScheme,localTuning}.
 
@@
 
-% Head 2
 \subsection{DIMENSION REDUCTION}
-There is a number of ways to adapt effective one-dimensional algorithms for solving multidimensional problems; see, for example, the diagonal partitions method in [15,16] or the simplicial partitions method in [17].
+There is a number of ways to adapt effective one-dimensional algorithms for solving multidimensional problems; see, for example, the diagonal partitions method in \citep{sergKvaDiagNew,kvasovDiag} or the simplicial partitions method in \citep{DISIMPL}.
 
@@
 $u=u_1=y$ , then solving the problem is an equivalent to its solving using a single evolvent mapping $[0,1]$ onto $D$; the nested subproblems are absent.
-The issues of the numerical construction of the Peano curve approximations and of using the nested optimization scheme are considered in details in [3,4].
+The issues of the numerical construction of the Peano curve approximations and of using the nested optimization scheme are considered in details in \citep{recogAppl,determApproaches}.
 
@@
 
-A detailed description of various variants of this algorithm is presented in [5, 18, 19, 20].
+A detailed description of various variants of this algorithm is presented in \cite{strongSerg,gergelDerivMethod,branchBoundAlg,barkAdaptConstr}.
 
@@
 
-Various variants of the parallel algorithms based on this idea are presented in [7-12].
+Various variants of the parallel algorithms based on this idea are presented in \citep{parAlg,parAlg2,gergelLNCS,gergelStrParComputing,mixedAlgLNCS,barkXeon}.
 
@@
 The idea of acceleration is to magnify the characteristics of the intervals containing the best current estimates by introducing some factor depending on the function values estimated at the end points of the corresponding interval.
-Following this idea, the “local” characteristics of the intervals can be introduced [5, 21].
+Following this idea, the “local” characteristics of the intervals can be introduced \citep{strongSerg,indexMethod}.
 
@@
 
-The problems generated by GKLS generator [22] were used as the test ones.
+The problems generated by GKLS generator \citep{GKLS} were used as the test ones.
 This generator allows generating the multiextremal optimization problems with the properties known in advance: number of local minima, size of attraction domains, global minimum point, function values, etc.

### 2016-05-10 2016/LION10/GenOpt.tex
Add first version of LION10 paper

\documentclass{llncs}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage[misc,geometry]{ifsym}

%\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage{epstopdf}
%\usepackage{epsfig}

%\usepackage{url}
%\urldef{\mailsa}\path|i.surmin@gmail.com, avbk@mail.ru, bastrakov@vmk.unn.ru,  |
%\urldef{\mailsb}\path|evgeny.efimenko@gmail.com, arkady.gonoskov@gmail.com, meerov@vmk.unn.ru|

\begin{document}

\mainmatter

\title{Solving GENOPT problems \newline with the use of ExaMin solver}
\author{Konstantin Barkalov (\email{konstantin.barkalov@itmm.unn.ru}[\Letter]) \and Alexander Sysoyev (\email{sysoyev@vmk.unn.ru})  \and Ilya Lebedev \email{lebedev.ilya.g@gmail.com } \and Vladislav Sovrasov (\email{sovrasov.vlad@gmail.com}) }

\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia}

\maketitle

\begin{abstract}

This paper describes an algorithm for solving multidimensional multiextremal optimization problems. This algorithm uses Peano-type space-filling curves for dimension reduction. It has been used for solving problems at GENeralization-based
contest in global OPTimization (GENOPT). Computational experiments are carried out on 1800 multidimensional problems.

\keywords global optimization, multiextremal functions, space-filling curves, mixed global-local algorithm, GENOPT.

\end{abstract}

\section{Introduction}
A well-known approach to the investigation and comparing of the multiextremal optimization algorithms is based on testing these methods by solving a set of test problems, chosen randomly from some specially designed class. Each test problem can be considered as a particular instance of a random function generated by a special generator. Application of multiextremal optimization algorithms to large sets of such functions allows estimating the characteristics of the methods and evaluating the efficiency of each particular algorithm.

Among such generators for the one-dimensional problems, there are samples from Fourier series proposed by Hill \cite{Hill}. A generator proposed by Shekel \cite{Shekel} generates another well-known class of the one-dimensional test problems. For the investigation of various one-dimensional algorithms using the samples of the functions generated by Hill and Shekel generators Globalizer software has been developed. A comprehensive description of the capabilities of this system and the examples of its application are given in \cite{Strongin2000}. Note also that Hill and Shekel functions have been successfully used in the construction of one-dimensional constrained problems (with controlled measure of the feasible domain) \cite{Barkalov2002}.

A generator for a random sampling of two-dimensional test functions has been proposed in \cite{Grishagin1978}. A generator for the functions of arbitrary dimensionality with known positions of the local and global minima (GKLS generator) has been proposed in \cite{Gaviano}. Its application for the studying of some multidimensional algorithms has been described in \cite{Kvasov2003,Sergeyev2013}.

This approach to the comparison of the algorithms has been used by the organizers of GENeralization-based contest in global OPTimization (GENOPT). The functions to be optimized are broadly divided into three function families: GKLS, conditioned transforms of classical benchmarks (Rosenbrock (unimodal, narrowing bending valley), Rastrigin (strongly multimodal), Zakharov (unimodal)), and a composition of classical benchmarks. In their turn, each family is subdivided into six function types, which differ by their analytical definition and by the number of dimensions. The functions belonging to the same type share the majority of properties and can be assumed to have the same difficulty. Finally, every function type of every family is realized into an unlimited number of function instances which differ by some randomly generated parameters. In particular, every instance will have a randomly generated offset $c\in[-1,1]$ added to the function’s output in order to make the global minimum value unpredictable. A detailed description of the competition problems can be found at the website \url{http://genopt.org}.

An efficient approach for solving global optimization problems has been developed under the supervision by prof. R.G. Strongin at Lobachevsky State University of Nizhni Novgorod \cite{Sergeyev1994} -- \cite{Barkalov2016}. Within the framework of this approach, solving multidimensional global optimization problems is reduced to solving a set of the corresponding one-dimensional ones. The corresponding dimension reduction is based on the use of Peano space-filling curves (also called \textit{evolvents}), unambiguously mapping the unit interval of the real axis onto a hypercube, as well as the generalization of these ones, which can be applied to solving the problems using the multiprocessor systems. The proposed algorithms have been implemented in ExaMin solver applied to solving the problems within GENOPT competition. In the present work, a brief description of the global optimization algorithm applied and of its modifications are given, and the results of the computational experiments with the problems of the competition are presented.

\section{Problem statement}

Let us consider the problem of search for a global minimum of an $N$-dimensional function $\varphi(y)$ within a hyperinterval $D$
\begin{eqnarray}\label{eq:1}
& \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
& D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
\end{eqnarray}

Let us assume that the function $\varphi$ satisfies the Lipschitz condition with an a priori unknown constant $L$
\[
\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D,\; 0<L<\infty.
\]

In this study, we will use the approach based on the idea of dimension reduction by means of a Peano curve $y(x)$, which continuously and unambiguously maps the unit interval [0,1] onto the $n$-dimensional cube
\[
\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq N\right\}=\left\{y(x):0\leq x \leq 1 \right\}.
\]
%new text begin
Problems of numerical construction of Peano-type space filling curves and the corresponding theory are considered in detail in \cite{Strongin2000}, \cite{Sergeyev2013}.  
%new text begin
Here we will note that a numerically constructed curve is $2^{-m}$ accurate approximation of the theoretical Peano curve in $L_\infty$ metric, where $m$ is an evolvent construction parameter. Examples of the evolvent with different $m$ in two dimensions are given in Fig.~\ref{fig:0}.

\begin{figure}
\begin{minipage}{0.32\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1a.jpg} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}{0.32\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1b.jpg} \\ (b)}
\end{minipage}
\begin{minipage}{0.32\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1c.jpg} \\ (c)}
\end{minipage}
\caption{Evolvents in two dimensions with (a) $m=3$, (b) $m=4$ and (c) $m=5$}
\label{fig:0}
\end{figure}

By using this kind of mapping it is possible to reduce the multidimensional problem~(\ref{eq:1}) to a univariate problem
\[
\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}}.
\]
An important property of such mapping is preservation of boundedness of function relative differences  (see \cite{Strongin2000,Sergeyev2010}): if the function $\varphi(y)$ in the domain $D$ satisfies the Lipschitz condition, then the function $\varphi(y(x))$ on the interval $[0,1]$ will satisfy a uniform H{\"o}lder condition
\[
\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N},
\]
where the H{\"o}lder constant $H$ is linked to the Lipschitz constant $L$ by the relation
\[
H=2L\sqrt{N+3}.%,\;\; d=\max{\left\{b_i-a_i:1\leq i \leq N\right\}}.
\]

Therefore, it is possible, without loss of generality, to consider minimization of univariate function
\[
f(x)=\varphi(y(x)), \;\; x\in[0,1],
\]
satisfying the H{\"o}lder condition.

\section{Global search algorithm}\label{sec:1}

The considered algorithm for solving this problem (here, according to \cite{Strongin2000}) involves constructing a sequence of points $x^k$, where the values of minimized function $z^k = f(x^k)=\varphi(y(x^k))$ are calculated. Let us call the process of calculating the function value (including the construction of an image $y^k=y(x^k)$) the ``trial'', and the pair $(x^k, z^k)$, the ``trial result''. The set of pairs $\left\{(x^k, z^k)\right\}, 1\leq k\leq n,$ makes up the search data collected using the method after carrying out $n$ steps. The rules that determine the work of the global search algorithm are as follows.

At the first iteration of the method the trial is carried out at an arbitrary internal point $x^1$ of the interval $[0,1]$. The point of trial at the next iteration $(k+1)$ is determined according to the rules presented below.

Rule 1. Renumber points of the set
\[
X_k=\{x^1,\dots,x^k\}\cup\left\{0\right\}\cup\left\{1\right\},
\]
which includes boundary points of the interval $[0,1]$ and the points of the previous trials, with subscripts in increasing order of coordinate values, i.e.,
\[
0=x_0<x_1<\dots <x_k<x_{k+1}=1.
\]

Rule 2. Supposing that  $z_i=f(x_i)=\varphi(y(x_i)), \; 1\leq i \leq k$, calculate values 
\begin{equation}\label{eq:11}
\mu = \max_{2\leq i \leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
\end{equation}
\[
M = \left\{
   \begin{array}{lr}
     r\mu, & \mu > 0,\\
     1, & \mu = 0,
   \end{array}
\right.
 \]
where $r>1$ is a preset parameter of the method, and $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$.

Rule 3. Calculate a \textit{characteristic} for every interval $(x_{i-1}, x_i), \; 1\leq i \leq k+1,$   according to the following formulae
\[
R(1)=2\Delta_1-4\frac{z_1}{M},
\]
\begin{equation}\label{eq:14}
R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\frac{z_i+z_{i-1}}{M},1<i<k+1,
\end{equation}
\[
R(k+1)=2\Delta_{k+1}-4\frac{z_k}{M}.
\]

Rule 4. Find interval $(x_{t-1},x_t)$ with the maximum characteristic
\begin{equation}\label{eq:141}
R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
\end{equation}

Rule 5. Carry out a trial at the point $x^{k+1}\in(x_{t-1},x_t)$, calculated using the following formulae
\[
x^{k+1} = \frac{x_t+x_{t-1}}{2}, \textrm{ if } t=1 \textrm{ or } t=k+1,
\]
\begin{equation}\label{eq:142}
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N, \textrm{ if } 1<t<k+1.
\end{equation}

The algorithm terminates if the condition $\Delta_t<\epsilon$ is satisfied; here $\epsilon>0$ is the preset accuracy. As the \textit{current estimate} of an optimum at the step $k$ we accept the value 
\begin{equation}\label{eq:phistar}
\varphi_k^\ast=\min_{1\leq i \leq k}z^i,
\end{equation}
and the vector 
\[
y_k^\ast=\arg \min_{y \in \{ y^1,...,y^k\} }\varphi(y).
\]

Let us illustrate the work of this \textit{global search algorithm} (GSA) for minimization of a multiextremal function of two variables generated by GKLS generator \cite{Gaviano}. %function No 4, simple
The experiment used the method parameters $r=3$, $\epsilon=10^{-3}$, and the evolvent construction parameter $m=10$. Fig.~\ref{fig:1} shows the level lines of the function and the points of 421 trials carried out by the method before the required accuracy was obtained.
\begin{figure}
	\center
  \includegraphics[width=0.75\textwidth]{fig2.jpg} 
  \caption{Minimization of a GKLS function of two variables by GSA}
  \label{fig:1}       % Give a unique label
\end{figure}

The following theorem from \cite{Strongin2000} determines sufficient convergence conditions of the global search algorithm.

\textbf{Theorem}. Let the point $\overline{y}$ be the limit point of the sequence ${y^k}$ generated by the rules of GSA while minimizing the Lipschitzian with the constant $L$ function $\varphi(y), y \in D$. Then:
\begin{enumerate}
	\item If side by side with $\overline{y}$ there exists another limit point $y'$ of the sequence ${y^k}$ , then $\varphi(\overline{y})=\varphi(y')$.
	\item For any $k \geq 1$ $x^k=\varphi (y^k) \geq \varphi (\overline{y})$.
	\item If at some step of the search process the value $\mu$ from (\ref{eq:11}) satisfies the condition 
\[
  r\mu > 2^{3-1/N}L\sqrt{N+3},
\]
than $\overline{y}$ is a global minimizer of the function $\varphi (y)$ over $D$ and any global minimizer $y^\ast$ from () is also a limit point of the sequence ${y^k}$.

\end{enumerate}

\textbf{Remark}. As follows from the theorem, the limit points of the trial sequence ${y^k}$ generated by GSA are global minimizers. This property makes GSA substantially different from search techniques one way or another based on the random search concept which generates trial sequences everywhere dense in the search domain.

\section{Tuning the method for GENOPT problems}\label{sec:3}

The organizers of GENOPT competition have offered 18 classes of problems ($n=0,1,2$, see table \ref{tab:problems}). In each class, 100 functions have been given.

\begin{table}
	\caption{The classes of problems to be solved within the competition}
	\label{tab:problems}
	\center
	\begin{tabular}{cccc}
		\hline\noalign{\smallskip}
		Index & Family & Type & Dimension \\
		\hline\noalign{\smallskip}
		0 &	 \multirow{6}{*}{GKLS} &	\multirow{2}{*}{non-differentiable} & 10 \\
		1 &	 &	 & 30 \\
		2 &	  &	\multirow{2}{*}{differentiable} & 10 \\
		3 &	 &	 & 30 \\
		4 &	  &	\multirow{2}{*}{twice differentiable} & 10 \\
		5 &	 &	 & 30 \\
		\hline\noalign{\smallskip}		
		6 &	\multirow{6}{*}{High condition}  &	\multirow{2}{*}{Rosenbrock} & 10 \\
		7 &	 &	 & 30 \\
		8 &	  &	\multirow{2}{*}{Rastrigin} & 10 \\
		9 &	 &	 & 30 \\
		10 &	  &	\multirow{2}{*}{Zakharov} & 10 \\
		11 &	 &	 & 30 \\
		\hline\noalign{\smallskip}
		$12+2n$ &	\multirow{2}{*}{Composite}  &	 & 10 \\
		$13+2n$ &	 &	 & 30 \\
\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

The limit of 1 million computations of the function (trials) imposed by the organizers is essential for finding the global minimum of the functions with given dimensionalities. Thus, even for the 10-dimensional problems, having built a uniform grid with four points in each dimension, one goes already out of the 1 million points limit. The efficient algorithms, building the grids not dense everywhere, place the trials much more efficiently. However, for the 30-dimensional problems such an efficiency appears to be not enough as well. In this section, we describe four modifications of the core method used in solving various problems within the competition. The first two modifications have been applied to all problems, the third one – to the problems of GKLS class, and the last one has been used for solving the problems with Rastrigin, Rosenbrock, and Zakharov functions.

\subsection{Mixed global-local algorithm}\label{mixed}

One scheme aiming to accelerate the search process is to be outlined in this subsection (other schemes described in \cite{Kvasov2006,Kvasov2015} ). The idea of acceleration is to magnify the characteristics of the intervals containing the best current estimates by introducing some factor depending on the function values estimated at the end-points of the corresponding interval. Following this idea, we introduce the modified characteristics 
\begin{equation}\label{eq:localR}
  R_\alpha(i) = \frac{R(i)}{\sqrt{(z_i-\varphi_k^\ast)(z_{i-1}-\varphi_k^\ast)}+\mu (1.5)^{-\alpha}},
\end{equation}
where $\mu$, $R(i)$ and $\varphi_k^\ast$ are respectively from (\ref{eq:11}), (\ref{eq:14}), (\ref{eq:phistar}), and $\alpha$ is some integer parameter for setting the desired level of localization.

Next, it is possible to build a scheme \textit{mixing} the 'global' and the 'local' decision rules, i.e. switching between formulae (\ref{eq:14}) and (\ref{eq:localR}) in some systematic way. In our experiments we use \textit{global-to-local} ratio $q$, specifying the number of global trials preceding each local trial.

The mixed strategy has the following features. 	Firstly, both decision rules are based on the same information, so that each decision action (no matter local or global) uses the outcomes of all the trials performed. Secondly, non-stop global search assures the global convergence; the aim of the local refinement is to accelerate the attainment of low function values.

Let us use a mixed global-local algorithm for solving a problem from section \ref{sec:1} with the global-to-local ratio $q=4$ and level of localization $\alpha = 15$. With this parameters the number of iterations was $203$. Fig.~\ref{fig:2} shows level lines of the same function from Fig.~\ref{fig:1} with points of the trials performed by the mixed method.

\begin{figure}
	\center
  \includegraphics[width=0.75\textwidth]{fig3.jpg} 
  \caption{Minimization of a GKLS function of two variables by mixed GSA}
  \label{fig:2}       % Give a unique label
\end{figure}

\subsection{Local refinement of the best current estimate} \label{hooke}

The second implemented modification of the core method consisted in a direct utilizing of a local optimization method, namely Hooke--Jeeves method \cite{HookeJeeves} (see also \cite{Wilde}, \cite{Himmelblau}). Schematically, the combined method works as follows:

\begin{itemize}
	\item The global phase

	\begin{itemize}
		\item To perform the GSA iterations until the current optimum (the minimal value of the objective function in the trial points computed already) is renovated.
	\end{itemize}

	\item The local phase
		\begin{itemize}
		\item Start Hooke-Jeeves method from the current optimum point with given exit condition according to precision.
		\item Add all the trial points of the local method into the GSA trial points database.
		\item Upon achieving given accuracy by the local method, go to the global phase.
	\end{itemize}

\end{itemize}

The accuracy of the local method (the termination condition) was taken as $10^{-5}$ for all problem classes of GKLS as well as for the Rastrigin and composite ones,  $10^{-7}$ for Rosenbrock functions, and $10^{-6}$ for Zakharov ones.

\subsection{Utilizing the random search for GKLS-30}\label{random}

The modification of the core method considered above have allowed solving almost all problems from the 10-dimensional GKLS classes (see more details in section \ref{sec:5}). However, these ones appeared to be insufficient for the 30-dimensional problems of this class. 
The multistart scheme was the next variant of the modification. Using Sobol quasi-random number generator, we have selected 500 start points. A local optimization method with the limitation not only in precision (stated above) but also in the number of iterations (1000) has been started in each point. Thus, 500 thousand trials could be spent in the local phase in the worst case. All the points of the trials executed at the local phase are stored in the database of the global method. Then, the global phase was started, when the rest of 1 million trials were executed using GSA.

\subsection{Use of separable search as an initial stage}\label{separable}

Since Rastrigin function is a separable one, one can search for its global optimum by performing the optimization in each coordinate separately. Thus, the following scheme has been employed in order to solve Rastrigin problem:

\begin{itemize}
	\item The separable stage 
	\begin{itemize}
		\item Select a start point.
		\item For each coordinate to perform:
		\begin{itemize}
			\item Fix all coordinated except the current one.
			\item Perform the optimization by one-dimensional GSA. 
		\end{itemize}
		\item Store all the trial points of the local method into the trial point database of GSA.		
	\end{itemize}
	\item The local stage 
	\begin{itemize}
		\item Start Hooke-Jeeves method with given exit condition according to precision from the point of current optimum.
		\item Store all the trial points of the local method into the trial point database of GSA.
		\item Upon achievement given precision by the local method, go to the global stage.
	\end{itemize}
\item The global stage
	\begin{itemize}
		\item Perform the iterations of GSA, until the current optimum (the minimum value of the objective function in the trial points computed already) is renovated.
	\end{itemize}	
\end{itemize}

The scheme considered above has allowed us to solve all the problems with 10-dimensional Rastrigin functions as well as with the 30-dimensional ones.

The same approach has been applied to solving the problems based on the unimodal functions (Rosenbrock and Zakharov ones) as well. In these cases, the separable stage provides a good initial approximation for the local method. Without the use of the separability, the local method starts from a point located far away from the global minimum and performs too many iterations until the termination conditions are satisfied.

\section{Numerical experiments}\label{sec:5}

The methods considered in sections \ref{sec:1} and \ref{sec:3} and the modifications of these ones have been implemented in ExaMin solver intended for the parallel solving of the multidimensional multiextremal global optimization problems developed in Lobachevsky State University of Nizhni Novgorod. Global search algorithm and block nested optimization scheme \cite{Barkalov2014} make the algorithmic basis for ExaMin solver. According to the competition conditions, the sequential mode of the solver execution only has been used when solving the problems. However, ExaMin supports the systems with distributed memory (using MPI) as well as with shared memory (using OpenMP). Moreover, NVIDIA graphic processors and Intel Xeon Phi coprocessors are supported.

In the final stage of the competition, ExaMin solver has taken the 3rd prize in the overall ranking and the 1st one in the total number of the solved problems (fig.~\ref{fig:4}). The distribution of the classes of the solved tasks is presented in table \ref{tab:solved}. The parameter of the evolvents building was $m=10$. The parameter of the method was $r=2.5$ for all problems besides the ones from GKLS class. For GKLS, $r$ varied from 2.5 to 20.

\begin{figure}
	\center
  \includegraphics[width=1.00\textwidth]{fig4.jpg} 
  \caption{Final GENOPT leaderboard}
  \label{fig:4}       % Give a unique label
\end{figure}

\begin{table}
	\caption{Total number of solved tasks from different classes}
	\label{tab:solved}
	\center
	\begin{tabular}{cc}
		\hline\noalign{\smallskip}
	Class & Tasks solved   \\
	\noalign{\smallskip} \hline \noalign{\smallskip}		
		GKLS--nd--10 & 99 \\
		GKLS--nd--30 & 15 \\
		GKLS--cd--10 & 96 \\
		GKLS--cd--30 & 1 \\
		GKLS--td--10 & 94 \\
		GKLS--td--30 & 0 \\
		Rosenbrock--10 & 100 \\
		Rosenbrock--30 & 100 \\
		Rastrigin--10 & 100 \\
		Rastrigin--30 & 100 \\
		Zakharov--10 & 100 \\
		Zakharov--30 & 100 \\
		Composite--10 & 100 \\
		Composite--30 & 100 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

The modification with the separable stage with the precision $0.02$ for the stop condition of one-dimensional GSA was used for all classes except GKLS. When solving the GKLS problems, the modifications from subsections \ref{mixed} and \ref{hooke} have been used first, then modification from subsection \ref{random} was applied. The distribution of the number of the solved tasks with the use of random search and without the one is presented in table \ref{tab:randsolved}.

\begin{table}
	\caption{Total number of solved tasks from different classes}
	\label{tab:randsolved}
	\center
	\begin{tabular}{ccc}
		\hline\noalign{\smallskip}
	Class &  \parbox[c]{3cm}{ \centering  Solved without\\random mod}   & \parbox[c]{3cm}{ \centering Solved with\\random mod}    \\
	\noalign{\smallskip} \hline \noalign{\smallskip}		
		GKLS--nd--10 & 78 & 99 \\
		GKLS--nd--30 & 0 & 15 \\
		GKLS--cd--10 & 67 & 96 \\
		GKLS--cd--30 & 0 & 1 \\
		GKLS--td--10 & 65 & 94 \\
		GKLS--td--30 & 0 & 0 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\section{Conclusion}
In this work, the results of solving of the 10- and 30-dimensional problems of the unconditional global optimization from GenOpt 2016 competition are presented. The optimization methods used and the modifications of these ones directed onto the obtaining of a solution at given limitation of 1 million trials (computations of the objective function) are described. All the modifications considered are implemented in ExaMin solver employed in the conducting of the experiments. The numerical experiments have been carried out using Lobachevsky supercomputer \cite{Lobach}.\\

%\subsection{Acknowledgements}

\textbf{Acknowledgements.}This study was supported by the Russian Science Foundation, project No 15-11-30022 ``Global optimization, supercomputing computations, and applications''.

\begin{thebibliography}{99}
\bibitem{Hill}
Hill, J.D.: A search technique for multimodal surfaces. IEEE Transactions on Systems Science and Cybernetics. 5(1), 2--8 (1969)

\bibitem{Shekel}
Shekel J.: Test functions for multimodal search technique. Proceedings of the 5th Princeton Conference on Information Science Systems. Princeton, Princeton University Press. 354--359 (1971)

\bibitem{Strongin2000}
Strongin, R.G., Sergeyev, Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000)

\bibitem{Barkalov2002}
Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Computational Mathematics and Mathematical Physics 42(9), 1289--1300 (2002)

\bibitem{Grishagin1978}
Grishagin, V.A.: Operation Characteristics of Some Global Optimization Algorithms. Problems
of Stochastic Search 7, 198--206 (1978) (in Russian)

\bibitem{Gaviano}
Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Transactions on Mathematical Software 29(4), 469--480 (2003)

\bibitem{Kvasov2003}
Kvasov, D., Sergeyev, Ya.: Multidimensional global optimization algorithm based on adaptive diagonal curves. Computational Mathematics and Mathematical Physics 43(1), 40--56 (2003)

\bibitem{Sergeyev2013}
Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013)

\bibitem{Sergeyev2010}
Lera D., Sergeyev Ya.D. Lipschitz and Holder global optimization using space-filling curves, Applied Numerical Mathematics, 60(1-2), 115--129 (2010)

\bibitem{Sergeyev1994}
Sergeyev, Ya.D., Grishagin, V.A.: A parallel method for finding the global minimum of univariate functions. J. Optim. Theory Appl. 80(3), 513--536 (1994)

\bibitem{SergeyevGrishagin}
Sergeyev, Ya.D., Grishagin, V.A.:  Sequential and parallel global optimization algorithms. Optimization Methods and Software. 3, 111--124 (1994)

\bibitem{Gergel1996}
Gergel, V.P.: A method of using derivatives in the minimization of multiextremum functions. Computational Mathematics and Mathematical Physics 36(6), 729--742 (1996)

\bibitem{Gergel1997}
Gergel, V.P.: A global optimization algorithm for multivariate functions with lipschitzian first derivatives. J. Glob. Optim. 10(3), 257--281 (1997)

\bibitem{Grishagin1997}
Grishagin, V.A., Sergeyev, Ya.D., Strongin, R.G.: Parallel characteristical algorithms for solving problems of global optimization. J. Glob. Optim. 10(2), 185--206 (1997)

\bibitem{Gergel1999}
Gergel, V.P., Sergeyev, Ya.D.: Sequential and parallel algorithms for global minimizing functions with Lipschitzian derivatives. Computers and Mathematics with Applications, 37(4-5), 163--179 (1999)

\bibitem{Grishagin2001}
Sergeyev Ya.D, Grishagin V.A.: Parallel asynchronous global search and the nested optimization scheme. Journal of Computational Analysis and Applications. 3(2), 123--145 (2001)

\bibitem{Strongin2003}
Strongin, R.G., Sergeyev, Ya.D.: Global optimization: fractal approach and non-redundant parallelism. J. Glob. Optim. 27(1), 25--50 (2003)

\bibitem{Gergel2005}
Gergel, V.P., Strongin, R.G.: Parallel computing for globally optimal decision making on cluster systems. Future Generation Computer Systems, 21(5), 673--678 (2005)

\bibitem{Barkalov2013}
Barkalov, K., Polovinkin, A., Meyerov, I., Sidorov, S., Zolotykh, N.: SVM regression parameters optimization using parallel global search algorithm. Lecture Notes in Computer Science. 7979, 154--166 (2013)

\bibitem{Barkalov2014} 
Barkalov, K.A., Gergel, V.P.: Multilevel scheme of dimensionality reduction for parallel global search algorithms. Proceedings of the 1st International Conference on Engineering and Applied Sciences Optimization -- OPT-i 2014. pp. 2111--2124. (2014)

\bibitem{Grishagin2015} 
Gergel, V., Grishagin, V., Israfilov, R.: Local tuning in nested scheme of global optimization. Procedia Computer Science, 51(1), pp. 865--874. (2015)

\bibitem{Gergel2015} 
Gergel, V., Grishagin, V., Gergel, A.: Adaptive nested optimization scheme for multidimensional global search. Journal of Global Optimization, 17 p. Article in Press. (2015)

\bibitem{Barkalov2016} 
Barkalov, K., Gergel, V.: Parallel global optimization on GPU. Journal of Global Optimization. 18 p. Article in Press. (2016)

\bibitem{Kvasov2006} 
Sergeyev, Ya.D., Kvasov D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM Journal on Optimization, 16(3), 910--937 (2006)

\bibitem{Kvasov2015} 
Paulavicius, R., Sergeyev, Y., Kvasov, D., Zilinskas, J. Globally-biased DISIMPL algorithm for expensive global optimization, Journal of Global Optimization, 59(2-3), 545--567 (2015).

\bibitem{HookeJeeves} 
Hooke, R., Jeeves, T.A.: "Direct search" solution of numerical and statistical problems // Journal of the ACM. 8(2), 212--229 (1961)

\bibitem{Wilde} 
Wilde, D.J.: Optimum Seeking Methods. Prentice-Hall, Engelwood Cliffs, New Jersey (1964)

\bibitem{Himmelblau} 
Himmelblau, D.M.: Applied Nonlinear Programming. McGraw-Hill, New York (1972)

\bibitem{Lobach}
\url{http://www.top500.org/system/178472}

\end{thebibliography}

\end{document}

### 2016-05-10 2016/LION10/GenOpt.tex
Add final version of LION10 paper

@@
 \title{Solving GENOPT problems \newline with the use of ExaMin solver}
-\author{Konstantin Barkalov (\email{konstantin.barkalov@itmm.unn.ru}[\Letter]) \and Alexander Sysoyev (\email{sysoyev@vmk.unn.ru})  \and Ilya Lebedev \email{lebedev.ilya.g@gmail.com } \and Vladislav Sovrasov (\email{sovrasov.vlad@gmail.com}) }
+\author{Konstantin Barkalov \Letter \and Alexander Sysoyev \and \\ Ilya Lebedev \and Vladislav Sovrasov \\
+\email{ \{konstantin.barkalov,alexander.sysoyev,ilya.lebedev,\\vladislav.sovrasov\}@itmm.unn.ru}}
+
+%\author{Konstantin Barkalov \Letter \and Alexander Sysoyev \and \newline Ilya Lebedev \and Vladislav Sovrasov \email{[konstantin.barkalov,alexander.sysoyev,ilya.lebedev,\newline vladislav.sovrasov]@itmm.unn.ru}}
+
+%\author{Konstantin Barkalov \Letter \and Alexander Sysoyev \and Ilya Lebedev \and Vladislav Sovrasov \email{[konstantin.barkalov,alexander.sysoyev,ilya.lebedev,\newline vladislav.sovrasov]@itmm.unn.ru}}
+
+%\author{Konstantin Barkalov ( \Letter~\email{konstantin.barkalov@itmm.unn.ru})  \and \\ Alexander Sysoyev (\email{alexander.sysoyev@itmm.unn.ru}) \and \\ Ilya Lebedev (\email{ilya.lebedev@itmm.unn.ru})  \and \\ Vladislav Sovrasov (\email{vladislav.sovrasov@itmm.unn.ru}) }
 
@@
 
-This approach to the comparison of the algorithms has been used by the organizers of GENeralization-based contest in global OPTimization (GENOPT).
+Approach to the comparison of the algorithms by solving a set of test problems has been used by the organizers of GENeralization-based contest in global OPTimization (GENOPT).
 The functions to be optimized are broadly divided into three function families: GKLS, conditioned transforms of classical benchmarks (Rosenbrock (unimodal, narrowing bending valley), Rastrigin (strongly multimodal), Zakharov (unimodal)), and a composition of classical benchmarks.
@@
 
-Let us illustrate the work of this \textit{global search algorithm} (GSA) for minimization of a multiextremal function of two variables generated by GKLS generator \cite{Gaviano}.
-%function No 4, simple
+%new
+This \textit{global search algorithm} (GSA) was developed in the framework of information - statistical approach (see \cite{Strongin2000}).
+From this point of view the normalized characteristics $R(i)$ from (\ref{eq:14}) can be considered as probabilities of locating the global minimum within the interval  $(x_{i-1},x_i), 1\leq i \leq k+1$.
+Thus, at each iteration a new trial point is selected inside the interval, which has the greatest probability of finding the global minimum.
+At the same time, use of running estimates (\ref{eq:11}) of Lipschitz constant $L$ allows us to apply GSA to the problems with a priori unknown values of $L$.
+
+Let us illustrate the work of GSA for minimization of a multiextremal function of two variables generated by GKLS generator \cite{Gaviano}. %function No 4, simple
 The experiment used the method parameters $r=3$, $\epsilon=10^{-3}$, and the evolvent construction parameter $m=10$.
@@
 
-The organizers of GENOPT competition have offered 18 classes of problems ($n=0,1,2$, see table \ref{tab:problems}). In each class, 100 functions have been given.
+The organizers of GENOPT competition have offered 18 classes of problems (see table \ref{tab:problems}, %new
+where last two rows correspond to six classes of composite functions with $n=0,1,2$). In each class, 100 functions have been given.
 
@@
 Thus, even for the 10-dimensional problems, having built a uniform grid with four points in each dimension, one goes already out of the 1 million points limit.
-The efficient algorithms, building the grids not dense everywhere, place the trials much more efficiently.
+%new
+The global optimization algorithms that build non-uniform grid in the search domain put the points more efficiently. %new
 However, for the 30-dimensional problems such an efficiency appears to be not enough as well.
@@
 
-One scheme aiming to accelerate the search process is to be outlined in this subsection (other schemes described in \cite{Kvasov2006,Kvasov2015} ).
+One scheme aiming to accelerate the search process is to be outlined in this subsection (other schemes are described, for example, in \cite{Kvasov2006,Kvasov2015} ).
 The idea of acceleration is to magnify the characteristics of the intervals containing the best current estimates by introducing some factor depending on the function values estimated at the end-points of the corresponding interval.
@@
 When solving the GKLS problems, the modifications from subsections \ref{mixed} and \ref{hooke} have been used first, then modification from subsection \ref{random} was applied.
-The distribution of the number of the solved tasks with the use of random search and without the one is presented in table \ref{tab:randsolved}.
+%new
+The number of the solved problems with and without random search is presented in table \ref{tab:randsolved}.%new
 
@@
 		\hline\noalign{\smallskip}
-	Class &  \parbox[c]{3cm}{ \centering  Solved without\\random mod}   & \parbox[c]{3cm}{ \centering Solved with\\random mod}    \\
+	Class &  \parbox[c]{3cm}{ \centering  Solved without\\random mode}   & \parbox[c]{3cm}{ \centering Solved with\\random mode}    \\
 	\noalign{\smallskip} \hline \noalign{\smallskip}		
@@
 All the modifications considered are implemented in ExaMin solver employed in the conducting of the experiments.
-The numerical experiments have been carried out using Lobachevsky supercomputer \cite{Lobach}.\\
+The numerical experiments have been carried out using Lobachevsky supercomputer \cite{Lobach}.
 
@@
 
-\textbf{Acknowledgements.}This study was supported by the Russian Science Foundation, project No 15-11-30022 ``Global optimization, supercomputing computations, and applications''.
+\textbf{Acknowledgements.} This study was supported by the Russian Science Foundation, project No 15-11-30022 ``Global optimization, supercomputing computations, and applications''.
 

### 2016-07-14 2016/YSC/ysc_paper.tex
Add missed files from last commit

% easychair.tex,v 3.1 2011/12/30
%
% Select appropriate paper format in your document class as
% instructed by your conference organizers. Only withtimes
% and notimes can be used in proceedings created by EasyChair
%
% The available formats are 'letterpaper' and 'a4paper' with
% the former being the default if omitted as in the example
% below.
%
\documentclass[procedia]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

% This provides the \BibTeX macro
\usepackage{doc}
\usepackage{makeidx}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{subfig}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% If you plan on including some algorithm specification, we recommend
% the below package. Read more details on the custom options of the
% package documentation.
%
% \usepackage{algorithm2e}

% Some of our commands for this guide.
%
\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\sign}{sign}
\newtheorem{theorem}{Theorem}

\def\procediaConference{99th Conference on Topics of
  Superb Significance (COOL 2014)}

%\makeindex

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Local Tuning in }

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair

\titlerunning{Local Tuning in \ldots}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes into the list
% defined using \institute
%
\author{
    Vladislav V. Sovrasov%\thanks{Designed and implemented the class style}
}

% Institutes for affiliations are also joined by \and,
\institute{
  State university of Nizhny Novgorod,
  Nizhny Novgorod, Russia\\
  \email{sovrasov.vlad@gmail.com}
 }
%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{Sovrasov}

\begin{document}

\maketitle

\keywords{advanced numerical methods, deterministic global optimization, speed-up of convergence, derivative-free algorithms}

\begin{abstract}
  In the present paper, one of the methods of accounting for the local information on
  the objective function in the global Lipschitz optimization problems has been considered.
  In the course of solving such problems, a problem of estimating Lipschitz constant of the
  optimized function arises. According to the classic scheme, such an estimate is a single one
  for the whole search domain. The method of accounting for the local properties is based on
  the building of the estimates of Lipschitz constants for the search subdomains and has been
  investigated earlier for the one-dimensional case. For solving the multidimensional problems,
  the dimension reduction is applied: a multidimensional optimization problem is reduced to
  a one-dimensional problem, which the objective function satisfy Hölder condition in.
  In this paper, the application of building the local estimates of Hölder constant in the
  reduced multidimensional optimization problems using the same scheme to the one used
  in the one-dimensional Lipschitzian problems is considered.
\end{abstract}

%\setcounter{tocdepth}{2}
%{\small
%\tableofcontents}

%\section{To mention}
%
%Processing in EasyChair - number of pages.
%
%Examples of how EasyChair processes papers. Caveats (replacement of EC
%class, errors).

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}

The one-dimensional characteristic algorithms represent one of the classes of the
optimization algorithms used widely \cite{}. A number of schemes of reduction of the
multidimensional problems to one or several one-dimensional ones, i. e. the dimension
reduction schemes have been developed \cite{stronginBook}, \cite{butzCurves}.
\par
In recent years, when building the global optimization algorithms, an increased attention
is paid to accounting for the local properties of the optimized function. In the
case of the characteristic schemes, this approach allows developing a mixed algorithm
\cite{mixedAlg} as well as using different estimates of Lipschitz constant in different search
domain \cite{sergLocalTuningFirst}. Originally, the latter approach has been proposed for the one-dimensional
problems, also the investigation of this one when using the nested optimization scheme
has been carried out \cite{nestedLocal}. In all these cases, this approach have demonstrated itself to
be an efficient one, having allowed accelerating the convergence of the characteristic
optimization methods essentially. In the present paper, a generalization of the method
for building of the local estimates of Lipschitz constants onto the case of Hölder
functions obtained when reducing the multidimensional optimization problems is considered.

%------------------------------------------------------------------------------
\section{Problem Statements}
\label{sect:problem}

One of the global optimization problem statements can be formulated as follows:
to find the global minimum of an \(N\)-dimensional function \(\varphi(y)\) in a hyperinterval
\(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
In order to obtain an estimate of the global minimum from a finite number of computations
of the function values, \(\varphi(y)\) is required to satisfy Lipschitz condition.
\begin{displaymath}
\label{task}
\phi(y^*)=\min\{\varphi(y):y\in D\}
\end{displaymath}
\begin{displaymath}
\label{lip}
|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
\end{displaymath}
\par
The use of the evolvents i. e. the curves filling the space is a classic scheme
for the dimension reduction for the global optimization algorithms \cite{strOptBook}.
\begin{displaymath}
\label{cube}
\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
\end{displaymath}
\par
Such a mapping allows reducing a problem stated in a multidimensional space to solving
a one-dimensional one at the expense of worsening its properties. In particular,
the one-dimensional function \(\varphi(y(x))\) is not a Lipschitz one but is a Hölder one:
\begin{displaymath}
\label{holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
\end{displaymath}
where Hölder constant \(H\) is related to Lipschitz constant \(L\) by the relation
\begin{displaymath}
H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}
\end{displaymath}
\par
Therefore, not limiting the generality, one can consider the minimization of the
one-dimensional function \(f(x)=\varphi(y(x))\), \(x\in[0;1]\) satisfying Hölder condition.
The issues of numerical building of the mapping like Peano curve and the corresponding
theory have been considered in details in \cite{strOptBook}. Here we would note that an evolvent
built numerically is an approximation to the theoretical Peano curve with the precision
of the order \(2^{-m}\) where \(m\) is the building parameter of the evolvent.

%------------------------------------------------------------------------------
\section{Description of the Algorithm}
\label{sect:algorithm}

The considered algorithm of solving the stated problem implies the building of
a sequence of points \(x_k\), which the values of the minimized function \(z_k = f(x_k)\)
are computed in. Let us call the process of computation of the function value
(including the building of an image \(y_k=y(x_k)\)) a trial, and the pair \((x_k,z_k)\) ---
the result of the trial. A set of the pairs \(\{(x_k,z_k)\}, 1\leqslant k\leqslant n\)
makes the search information accumulated by the method after executing \(n\) steps.
\par
At the first iteration of the method, a trial is executed in an arbitrary internal
point \(x_1\) of the interval \([0,1]\). Let us assume \(n \geqslant 1\) iterations
of the method to be completed, in the course of which the iterations in \(k = k(n)\)
points \(x_i, 1\leqslant i\leqslant k\) have been performed. Then, the point \(x^{k+1}\)
of the search trial of the next \((k+1)\)th iteration is determined in accordance with the rules:
\par
Step 1. To renumber the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\),
which includes the boundary points of the interval \([0,1]\) as well as the points of
preceding trials, by the lower indices in the order of increasing the values of coordinate i. e.
\begin{displaymath}
0=x_0<x_1<\dotsc<x_{k+1}=1
\end{displaymath}
\par
Step 2. Assuming \(z_i=f(x_i),1\leqslant i\leqslant k\), to compute the quantities
\begin{equation}
\label{step2}
\mu=\max_{1\leqslant i\leqslant k}\dfrac{|z_i-z_{i-1}|}{\Delta_i},
\begin{matrix}
    M =
    \left\{
    \begin{matrix}
    r\mu,\mu>0 \\
    1,\mu=0
    \end{matrix} \right.
    \end{matrix}
\end{equation}
where \(r\) is a predefined parameter of the method, and \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
\par
Step 3. For each interval \((x_{i-1},x_i),1\leqslant i\leqslant k+1\), to compute the
characteristics according to the formulae
\begin{equation}
\label{step3_1}
R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M}
\end{equation}
\begin{equation}
\label{step3_2}
R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1
\end{equation}
\par
Step 4. Select the interval \((x_{t-1}, x_t)\) such that
\begin{equation}
\label{step4}
t=\argmax_{1\leqslant i \leqslant k+1}R(i),
\end{equation}
i.e., the interval with the maximal characteristic.
\par
Step 5. To execute new trial in the point \(x_{k+1}\) computed according to the formulae
\begin{displaymath}
x_{k+1}=\dfrac{x_{t}+x_{t-1}}{2},t=1,t=k+1
\end{displaymath}
\begin{equation}
\label{step5}
x_{k+1}=\dfrac{x_{t}+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{\mu}\right]^N,1<t<k+1
\end{equation}
\par
The algorithm is terminated if the condition \(\Delta_{t}\leqslant \varepsilon\) is fulfilled;
here \(\varepsilon>0\) is a predefined precision. As an estimate of the global optimum solution of the problem the values
\begin{equation}
f_k^*=\min_{1\leqslant i \leqslant k}f(x_i), x_k^*=\argmin_{1\leqslant i \leqslant k}f(x_i)
\end{equation}
\par
are selected. The theoretical substantiation of this method is presented in \cite{strOptBook}.

%------------------------------------------------------------------------------
\section{Local Adaptive Estimate of Hölder Constant}
%\label{sect:algorithm}
At it is seen from the scheme of the algorithm, regardless to the local properties
of the optimized one-dimensional function, the same values of estimate of Hölder
constant is used to compute the characteristic of all intervals (\ref{step2}, \ref{step3_1}, \ref{step3_2}).
In \cite{sergLocalTuning}, it has been proposed to use different values of \(M\) for each interval,
and also the efficiency of this approach in the case of the optimization of the
one-dimensional functions satisfying Lipschitz condition has been shown. In \cite{nestedLocal},
the application of the adaptive estimates of Lipschitz constants in the multidimensional
nested optimization scheme has been considered. For each interval, the local estimate of the
constant is an additive convolution of the «global» and «local» ones (\(\gamma\) and \(\lambda\), correspondingly):
\begin{displaymath}
  \begin{array}{lr}
    \lambda_i=max\{H_{i-1},H_i,H_{i+1}\} \\
    H_i=\frac{|z_i-z_{i-1}|}{\Delta_i} \\
    H^k=max\{H_i:i=2,\dots ,k\} \\
    \gamma_i=H^k\frac{\Delta_i}{\Delta^{max}} \\
    \Delta^{max}=max\{\Delta_{i}:i=2,\dots ,k\}
  \end{array}
\end{displaymath}
\begin{equation}
\label{additiveConv}
M_i=r\cdot max\{H_i, \frac{1}{2}(\lambda_i+\gamma_i),\xi\}
\end{equation}
\(\xi\) is chosen small to prevent the case of the function being constant over search interval.
This variant of convolution does not depend on the parameter \(r\), however, the adaptive convolution of the kind:
\begin{equation}
\label{additiveAdaptiveConv}
M_i=r\cdot max\{H_i, \frac{\lambda_i}{r}+\frac{r-1}{r}\gamma_i,\xi\}
\end{equation}
\par
has been considered in \cite{sergLocalTuning} as well.
If it is known a priori that the optimized function has a complex shape with multiple
local minima, then the initial value of \(r\) is specified to be large enough that
results in the dominance of the «global» component \(\gamma\) in the adaptive convolution.

%------------------------------------------------------------------------------
\section{Conditions of Convergence}
%\label{sect:algorithm}
In \cite{sergLocalTuning}, a theorem on the convergence of the method in the case if the objective
function is Lipschitz one is given. However, as a rule, such statements are true in
Hölder metrics as well, therefore, presumably, the following theorem is true:

\begin{theorem}
Assume the objective function \(f(x)\) to  satisfy Hölder condition with finite
constant \(H > 0\), and let x be a limit point of \(\{x_k\}\) generated by the algorithm.
Then, the following assertions hold:
\begin{enumerate}
  \item If \(x\in(0;1)\), then convergence to \(x\) is a bilateral one i. e. there
  exist two infinite subsequences of \(\{x_k\}\) converging to \(x\): one from the
  left, the other from the right;
  \item \(f(x_k) \geqslant f(x)\) for all trial points \(x_k, k \geqslant 1\);
  \item If there exists another limit point \(x^* = x\), then \(f(x) = f(x^*)\);
  \item If the function \(f(x)\) has a finite number of local minima in \([0, 1]\),
  then the point \(x\) is a local optimum one;
  \item (Sufficient conditions for convergence to a global minimizer). Let \(x^*\)
  be a global minimizer of \(f(x)\). If there exists an iteration number \(k^*\)
  such that for all \(k > k^*\) the inequality
  \(M_j(k) > H_j(k)\) holds, where \(H_j(k)\) is Hölder constant for the interval
  \([x_{j(k)-1}, x_{j(k)}]\) containing \(x^*\), and \(M_{j(k)}\) is its estimate.
  Then, the set of limit points of the sequence \(\{xk\}\) coincides with the set
  of global minimizers of the function \(f(x)\).
\end{enumerate}
\end{theorem}

The proof of this theorem requires further theoretical studies. Within the frames of
the present work, it is not performed. The convergence has been established numerically only.

%------------------------------------------------------------------------------
\section{Results of Experiments}
\label{sect:experiments}

The experiments on the evaluation of the efficiency of the method with the local
adaptive estimate of Hölder constant have been carried out using the two-dimensional
classes of Grishagin problems (\(F_{GR}\)) \cite{grishaginClass} and GKLS Simple 2d ones \cite{gklsClass}. Each class
includes 100 multiextremal functions. In all experiments, the evolvent was built
with the density \(m = 12\), parameter \(\varepsilon\) in the termination criterion was equal \(10^{-3}\).
The parameter \(r\) was selected as low as possible, which given method solves all
problems of the class at. The search step for \(r\) was \(0.1\).
\par
For clarity of illustration of the advantages of the local adaptive scheme of estimating
the constant \(H\), let us consider a particular example of the results obtained by the method.
In Fig. \ref{fig:grish_isolines}, the contour plots of a function from (\(F_{GR}\)) class and the trial
points executed by the method with the global estimate of Hölder constant and with
the one estimated according to formula (\ref{additiveConv}) are shown. As it is
seen from the figures, the method with global estimate of the constant executes a
large number of trials in the nearness of the global minimizer (total 1086 trials
have been executed) before the termination condition is satisfied, whereas the method
with the local adaptive estimate converges much faster (total 385 trials have been executed).
The same situation takes place when optimizing a function from GKLS Simple 2d class (Fig. \ref{fig:gkls_isolines}).
The method with the global estimate of the constant had executed 2600 trials while
the method with the local adaptive estimate had executed 1190 trials only.
\begin{figure}[ht]
    \centering
    \subfloat[global \(H\) estimation]{{\includegraphics[width=0.45\textwidth]{images/gs_glob.png} }}
    \qquad
    \subfloat[local-adaptive \(H\) estimation]{{\includegraphics[width=0.45\textwidth]{images/gs_loc.png} }}
    \caption{The contour plots of a function from (\(F_{GR}\)) class}
    \label{fig:grish_isolines}
\end{figure}

\begin{figure}[ht]
    \centering
    \subfloat[global \(H\) estimation]{{\includegraphics[width=0.45\textwidth]{images/gkls_glob.png} }}
    \qquad
    \subfloat[local-adaptive \(H\) estimation]{{\includegraphics[width=0.45\textwidth]{images/gkls_loc.png} }}
    \caption{The contour plots of a function from GKLS Simple 2d class}
    \label{fig:gkls_isolines}
\end{figure}
\par
Next, let us go to the comparison of various variants of the method for the problem
classes considered. In order to evaluate the efficiency of an algorithm, we will use
the operating characteristics \cite{grishaginClass}, which is defined by a set of points on the \((K, P)\)
plane where \((K)\) is the averaged number of the search trials before the satisfaction
of the termination condition when minimizing a function from given class, and \((P)\)
is the statistical probability of finding the global extremum is found with given
precision to the moment of termination. If at given \((K)\) the operating characteristic
of a method goes higher than the one of another method, it means that at fixed costs
of the search the former method will find the solution with a greater statistical probability.
If some value of \((P)\) is fixed, and the characteristic of a method goes to the left from
the one of another method, the former method requires fewer resources to achieve the same reliability.
\par
\begin{figure}[ht]
  	\center
    \includegraphics[width=0.75\textwidth]{images/grishagin.png}
    \caption{Operating characteristics of the methods compared on \(F_{GR}\) problems class}
    \label{fig:grishh_op}
  \end{figure}
As it is seen from the operating characteristics in Fig. \ref{fig:grishh_op}, \ref{fig:gkls_op}, both methods with the
local adaptive estimate of Hölder constant have demonstrated an essential advantage.
However, both methods require setting a higher value of the parameter \(r\). If one compares the
methods with the adaptive convolution and  with the non-adaptive one (\ref{additiveConv})(\ref{additiveAdaptiveConv}), the advantage
of the latter is seen clearly, although it requires a higher value of the reliability
parameter \(r\) to solve the problems form GKLS class.
  \begin{figure}[ht]
  	\center
    \includegraphics[width=0.75\textwidth]{images/gkls-s.png}
    \caption{Operating characteristics of the methods compared on GKLS Simple 2d problems class}
    \label{fig:gkls_op}
  \end{figure}
%------------------------------------------------------------------------------
\section{Conclusions}
In the present paper, the application of the methods of accounting for the local
behavior of the optimized function proposed earlier for the one-dimensional problems
in the multidimensional multiextremal optimization methods has been considered. Taking
into account the local properties is expressed in the use of different estimates of Hölder
constant in different search domains. The efficiency of the considered approach has been
confirmed by solving the essentially multiextremal problems form two test classes.

%------------------------------------------------------------------------------
% Refs:
%
\label{sect:bib}
%\bibliographystyle{plain}
%\bibliographystyle{alpha}
\bibliographystyle{unsrt}
%\bibliographystyle{abbrv}
\bibliography{ysc_paper}

%------------------------------------------------------------------------------
% Index
%\printindex

%------------------------------------------------------------------------------
\end{document}

% EOF

### 2016-07-14 2016/YSC/ysc_paper.tex
Add title and links

@@
 %
-\title{Local Tuning in }
+\title{Local Tuning in Peano curves-based global optimization scheme}
 
@@
 
-\titlerunning{Local Tuning in \ldots}
+\titlerunning{Local Tuning in Peano curves-based \ldots}
 
@@
 The one-dimensional characteristic algorithms represent one of the classes of the
-optimization algorithms used widely \cite{}. A number of schemes of reduction of the
+optimization algorithms used widely \cite{optHandbook}. A number of schemes of reduction of the
 multidimensional problems to one or several one-dimensional ones, i. e. the dimension
@@
 the constant \(H\), let us consider a particular example of the results obtained by the method.
-In Fig. \ref{fig:grish_isolines}, the contour plots of a function from (\(F_{GR}\)) class and the trial
+In Fig. \ref{fig:grish_isolines}, the contour plots of a function from \(F_{GR}\) class and the trial
 points executed by the method with the global estimate of Hölder constant and with
@@
     \subfloat[local-adaptive \(H\) estimation]{{\includegraphics[width=0.45\textwidth]{images/gs_loc.png} }}
-    \caption{The contour plots of a function from (\(F_{GR}\)) class}
+    \caption{The contour plots of a function from \(F_{GR}\) class}
     \label{fig:grish_isolines}
@@
 \label{sect:bib}
-%\bibliographystyle{plain}
+\bibliographystyle{plain}
 %\bibliographystyle{alpha}
-\bibliographystyle{unsrt}
+%\bibliographystyle{unsrt}
 %\bibliographystyle{abbrv}

### 2016-07-15 2016/YSC/ysc_paper.tex
Add Acknowledgements section to YSC paper
Small fixes in YSC paper

@@
 %
-\title{Local Tuning in Peano curves-based global optimization scheme}
+\title{Local Tuning in Peano Curves-based Global Optimization Scheme}
 
@@
 
-\titlerunning{Local Tuning in Peano curves-based \ldots}
+\titlerunning{Local Tuning in Peano Curves-based \ldots}
 
@@
 \institute{
-  State university of Nizhny Novgorod,
+  State University of Nizhny Novgorod,
   Nizhny Novgorod, Russia\\
@@
 
-\keywords{advanced numerical methods, deterministic global optimization, speed-up of convergence, derivative-free algorithms}
+\keywords{advanced numerical methods, deterministic global optimization, speedup of convergence, derivative-free algorithms}
 
@@
   In the present paper, one of the methods of accounting for the local information on
-  the objective function in the global Lipschitz optimization problems has been considered.
+  the objective function in the Lipschitz global optimization problems has been considered.
   In the course of solving such problems, a problem of estimating Lipschitz constant of the
-  optimized function arises. According to the classic scheme, such an estimate is a single one
+  objective function arises. According to the classic scheme, such an estimate is a single one
   for the whole search domain. The method of accounting for the local properties is based on
@@
   investigated earlier for the one-dimensional case. For solving the multidimensional problems,
-  the dimension reduction is applied: a multidimensional optimization problem is reduced to
+  the dimension reduction is applied. A multidimensional optimization problem is reduced to
   a one-dimensional problem, which the objective function satisfy Hölder condition in.
@@
 
-The one-dimensional characteristic algorithms represent one of the classes of the
+The one-dimensional characteristical algorithms represent one of the classes of the
 optimization algorithms used widely \cite{optHandbook}. A number of schemes of reduction of the
-multidimensional problems to one or several one-dimensional ones, i. e. the dimension
-reduction schemes have been developed \cite{stronginBook}, \cite{butzCurves}.
+multidimensional problems to one or several one-dimensional ones, i. e. the \textit{dimension
+reduction} schemes have been developed \cite{stronginBook}, \cite{butzCurves}.
 \par
 In recent years, when building the global optimization algorithms, an increased attention
-is paid to accounting for the local properties of the optimized function. In the
+is paid to accounting for the local properties of the objective function. In the
 case of the characteristic schemes, this approach allows developing a mixed algorithm
@@
 has been carried out \cite{nestedLocal}. In all these cases, this approach have demonstrated itself to
-be an efficient one, having allowed accelerating the convergence of the characteristic
+be an efficient one, having allowed accelerating the convergence of the
 optimization methods essentially. In the present paper, a generalization of the method
@@
 %------------------------------------------------------------------------------
-\section{Problem Statements}
+\section{Problem Statement}
 \label{sect:problem}
 
-One of the global optimization problem statements can be formulated as follows:
+The global optimization problem can be formulated as follows:
 to find the global minimum of an \(N\)-dimensional function \(\varphi(y)\) in a hyperinterval
-\(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
+\(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\): \(\varphi(y^*)=\min\{\varphi(y):y\in D\}\).
 In order to obtain an estimate of the global minimum from a finite number of computations
 of the function values, \(\varphi(y)\) is required to satisfy Lipschitz condition.
-\begin{displaymath}
-\label{task}
-\phi(y^*)=\min\{\varphi(y):y\in D\}
-\end{displaymath}
 \begin{displaymath}
@@
 \par
-Step 1. To renumber the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\),
+Step 1. Renumber the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\),
 which includes the boundary points of the interval \([0,1]\) as well as the points of
@@
 \par
-Step 2. Assuming \(z_i=f(x_i),1\leqslant i\leqslant k\), to compute the quantities
+Step 2. Assuming \(z_i=f(x_i),1\leqslant i\leqslant k\), compute the quantities
 \begin{equation}
@@
 \par
-Step 3. For each interval \((x_{i-1},x_i),1\leqslant i\leqslant k+1\), to compute the
+Step 3. For each interval \((x_{i-1},x_i),1\leqslant i\leqslant k+1\), compute the
 characteristics according to the formulae
@@
 \par
-Step 5. To execute new trial in the point \(x_{k+1}\) computed according to the formulae
+Step 5. Execute new trial in the point \(x_{k+1}\) computed according to the formulae
 \begin{displaymath}
@@
 \end{equation}
-\par
 are selected. The theoretical substantiation of this method is presented in \cite{strOptBook}.
@@
 of the optimized one-dimensional function, the same values of estimate of Hölder
-constant is used to compute the characteristic of all intervals (\ref{step2}, \ref{step3_1}, \ref{step3_2}).
+constant (\ref{step2}) is used to compute the characteristic of all intervals (\ref{step3_1}), (\ref{step3_2}).
 In \cite{sergLocalTuning}, it has been proposed to use different values of \(M\) for each interval,
@@
 nested optimization scheme has been considered. For each interval, the local estimate of the
-constant is an additive convolution of the «global» and «local» ones (\(\gamma\) and \(\lambda\), correspondingly):
+constant is an additive convolution of the ''global'' and ''local'' ones (\(\gamma\) and \(\lambda\), correspondingly):
 \begin{displaymath}
   \begin{array}{lr}
-    \lambda_i=max\{H_{i-1},H_i,H_{i+1}\} \\
+    \lambda_i=\max\{H_{i-1},H_i,H_{i+1}\} \\
     H_i=\frac{|z_i-z_{i-1}|}{\Delta_i} \\
-    H^k=max\{H_i:i=2,\dots ,k\} \\
+    H^k=\max\{H_i:i=2,\dots ,k\} \\
     \gamma_i=H^k\frac{\Delta_i}{\Delta^{max}} \\
-    \Delta^{max}=max\{\Delta_{i}:i=2,\dots ,k\}
+    \Delta^{max}=\max\{\Delta_{i}:i=2,\dots ,k\}
   \end{array}
@@
 \label{additiveConv}
-M_i=r\cdot max\{H_i, \frac{1}{2}(\lambda_i+\gamma_i),\xi\}
+M_i=r\cdot \max\{H_i, \frac{1}{2}(\lambda_i+\gamma_i),\xi\}
 \end{equation}
 \(\xi\) is chosen small to prevent the case of the function being constant over search interval.
-This variant of convolution does not depend on the parameter \(r\), however, the adaptive convolution of the kind:
+This variant of convolution does not depend on the parameter \(r\), however, the adaptive convolution:
 \begin{equation}
 \label{additiveAdaptiveConv}
-M_i=r\cdot max\{H_i, \frac{\lambda_i}{r}+\frac{r-1}{r}\gamma_i,\xi\}
+M_i=r\cdot \max\{H_i, \frac{\lambda_i}{r}+\frac{r-1}{r}\gamma_i,\xi\}
 \end{equation}
-\par
 has been considered in \cite{sergLocalTuning} as well.
@@
 local minima, then the initial value of \(r\) is specified to be large enough that
-results in the dominance of the «global» component \(\gamma\) in the adaptive convolution.
+results in the dominance of the ''global'' component \(\gamma\) in the adaptive convolution.
 
@@
 adaptive estimate of Hölder constant have been carried out using the two-dimensional
-classes of Grishagin problems (\(F_{GR}\)) \cite{grishaginClass} and GKLS Simple 2d ones \cite{gklsClass}. Each class
+classes of Grishagin problems (\(F_{GR}\)) \cite{grishaginClass} and GKLS ones \cite{gklsClass}. Each class
 includes 100 multiextremal functions. In all experiments, the evolvent was built
 with the density \(m = 12\), parameter \(\varepsilon\) in the termination criterion was equal \(10^{-3}\).
-The parameter \(r\) was selected as low as possible, which given method solves all
-problems of the class at. The search step for \(r\) was \(0.1\).
+The parameter \(r\) was selected as low as possible, allowing to solve all the
+problems. The search step for \(r\) was \(0.1\).
 \par
@@
 the constant \(H\), let us consider a particular example of the results obtained by the method.
-In Fig. \ref{fig:grish_isolines}, the contour plots of a function from \(F_{GR}\) class and the trial
+In Fig. \ref{fig:grish_isolines}, the level lines of a function from \(F_{GR}\) class and the trial
 points executed by the method with the global estimate of Hölder constant and with
@@
 with the local adaptive estimate converges much faster (total 385 trials have been executed).
-The same situation takes place when optimizing a function from GKLS Simple 2d class (Fig. \ref{fig:gkls_isolines}).
+The same situation takes place when optimizing a function from GKLS class (Fig. \ref{fig:gkls_isolines}).
 The method with the global estimate of the constant had executed 2600 trials while
@@
     \subfloat[local-adaptive \(H\) estimation]{{\includegraphics[width=0.45\textwidth]{images/gs_loc.png} }}
-    \caption{The contour plots of a function from \(F_{GR}\) class}
+    \caption{The level lines of a function from \(F_{GR}\) class}
     \label{fig:grish_isolines}
@@
     \subfloat[local-adaptive \(H\) estimation]{{\includegraphics[width=0.45\textwidth]{images/gkls_loc.png} }}
-    \caption{The contour plots of a function from GKLS Simple 2d class}
+    \caption{The contour plots of a function from GKLS class}
     \label{fig:gkls_isolines}
@@
 the operating characteristics \cite{grishaginClass}, which is defined by a set of points on the \((K, P)\)
-plane where \((K)\) is the averaged number of the search trials before the satisfaction
-of the termination condition when minimizing a function from given class, and \((P)\)
-is the statistical probability of finding the global extremum is found with given
-precision to the moment of termination. If at given \((K)\) the operating characteristic
+plane where \(K\) is the averaged number of the search trials carried out before the satisfaction
+of the termination condition when minimizing a function from given class, and \(P\)
+is portion of problems solved successfully. If at given \(K\) the operating characteristic
 of a method goes higher than the one of another method, it means that at fixed costs
-of the search the former method will find the solution with a greater statistical probability.
-If some value of \((P)\) is fixed, and the characteristic of a method goes to the left from
+of the search the former method will find the solution with a greater probability.
+If some value of \(P\) is fixed, and the characteristic of a method goes to the left from
 the one of another method, the former method requires fewer resources to achieve the same reliability.
@@
     \includegraphics[width=0.75\textwidth]{images/gkls-s.png}
-    \caption{Operating characteristics of the methods compared on GKLS Simple 2d problems class}
+    \caption{Operating characteristics of the methods compared on GKLS problems class}
     \label{fig:gkls_op}
@@
 In the present paper, the application of the methods of accounting for the local
-behavior of the optimized function proposed earlier for the one-dimensional problems
-in the multidimensional multiextremal optimization methods has been considered. Taking
+properties of the objective function
+in the multidimensional multiextremal optimization problems has been considered. Taking
 into account the local properties is expressed in the use of different estimates of Hölder
 constant in different search domains. The efficiency of the considered approach has been
-confirmed by solving the essentially multiextremal problems form two test classes.
+confirmed by solving the series of multidimensional problems form two classes.
+
+%------------------------------------------------------------------------------
+\section{Acknowledgements}
+This research was supported by the Russian Science Foundation, project No 16-11-10150
+“Novel efficient methods and software tools for the time consuming decision making problems
+with using supercomputers of superior performance”.
 

### 2016-07-15 2016/YSC/ysc_paper.tex
Deleted unused header, update YSC paper

@@
 is paid to accounting for the local properties of the objective function. In the
-case of the characteristic schemes, this approach allows developing a mixed algorithm
+case of the characteristical schemes, this approach allows developing a mixed algorithm
 \cite{mixedAlg} as well as using different estimates of Lipschitz constant in different search
@@
 \par
-The use of the evolvents i. e. the curves filling the space is a classic scheme
+The use of the evolvents \(y(x)\) i. e. the curves filling the space is a classic scheme
 for the dimension reduction for the global optimization algorithms \cite{strOptBook}.
@@
   \([x_{j(k)-1}, x_{j(k)}]\) containing \(x^*\), and \(M_{j(k)}\) is its estimate.
-  Then, the set of limit points of the sequence \(\{xk\}\) coincides with the set
+  Then, the set of limit points of the sequence \(\{x_k\}\) coincides with the set
   of global minimizers of the function \(f(x)\).
@@
 \label{sect:bib}
-\bibliographystyle{plain}
+%\bibliographystyle{plain}
 %\bibliographystyle{alpha}
-%\bibliographystyle{unsrt}
+\bibliographystyle{unsrt}
 %\bibliographystyle{abbrv}

### 2016-07-20 2016/LION10/GenOpt.tex
Финальная версия статьи ни LION, которую я выложил (по просьбе оргкомитета конференции) в EasyChair

@@
 \begin{minipage}{0.32\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{fig1a.jpg} \\ (a)}
+\center{\includegraphics[width=1.0\linewidth]{fig1a.JPG} \\ (a)}
 \end{minipage}
@@
 \begin{minipage}{0.32\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{fig1b.jpg} \\ (b)}
+\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (b)}
 \end{minipage}
 \begin{minipage}{0.32\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{fig1c.jpg} \\ (c)}
+\center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (c)}
 \end{minipage}
@@
 	\begin{itemize}
-		\item To perform the GSA iterations until the current optimum (the minimal value of the objective function in the trial points computed already) is renovated.
+		\item Perform the GSA iterations until the current optimum (the minimal value of the objective function in the trial points computed already) is renovated.
 	\end{itemize}

### 2016-08-03 2016/GOM/paper.tex
Add initial version of GOM paper

% gOMSguide.tex
% v5.0 released July 2015

\documentclass{gOMS2e}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables

\theoremstyle{plain}% Theorem-like structures
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\theoremstyle{remark}
\newtheorem{remark}{Remark}

\begin{document}

%\jvol{00} \jnum{00} \jyear{2015} \jmonth{July}

%\articletype{GUIDE}

\title{GLOBALIZER: A NOVEL SUPERCOMPUTER SOFTWARE SYSTEM FOR SOLVING TIME-CONSUMING GLOBAL OPTIMIZATION PROBLEMS}

\author{
\name{V. P. Gergel\textsuperscript{a}$^{\ast}$\thanks{$^\ast$Corresponding author. Email: latex.helpdesk@tandf.co.uk}
and K. A. Barkalov\textsuperscript{b} and A. V. Sysoev}
\affil{\textsuperscript{a}Taylor \& Francis, 4 Park Square, Milton Park, Abingdon, UK;
\textsuperscript{b}Institut f\"{u}r Informatik, Albert-Ludwigs-Universit\"{a}t, Freiburg, Germany}
\received{v5.0 released July 2015}
}

\maketitle

\begin{abstract}
In this paper, we describe the Globalizer software system for solving global optimization
problems. The system is designed to maximize the use of computational potential by modern
high performance computer systems in order to solve the most time-consuming optimization
problems. Highly parallel computations are facilitated using various distinctive computational
schemes: processing several optimization iterations simultaneously, reducing multidimensional
optimization problems using multiple Peano curve evolvents, multi-stage computing based on
nested block reduction schemes. This novelty leverages supercomputer system capabilities
with shared and distributed memory and with large numbers of processors to efficiently
solve global optimization problems.
\end{abstract}

\begin{keywords}
global optimization, information-statistical theory, parallel computing, high-performance
computer systems, supercomputer technologies
\end{keywords}

\section{Introduction}
Global (or multiextremal) optimization problems are among the most complex problems
in both theory and practice for optimal decision making. In these kinds of problems,
the criterion to be optimized has several local optima within the search domain,
which have different values. The existence of several local optima essentially makes
finding the global optimum difficult, since it requires examining the whole feasible
search domain. The volume of computations for solving global optimization problems
can increase exponentially with increasing number of varied parameters.
\par
In addition, global optimization problem statements are used, as a rule, in the most
complex decision making situations, for example, when a computer-aided design of complex
technologies, products, and systems is conducted. In such problems, the efficiency
criteria are nonlinear, the search domains may be non-contiguous, and, most importantly,
the computational complexity of the functions which the optimized criteria and constraints
are based on can be quite essential.
\par
These global optimization problem features impose special requirements on the quality of
the optimization methods and on the software to implement them. The global optimization
methods should be highly efficient, and the software systems should be developed on a good
professional basis. In general, global optimization problems can be solved using modern
supercomputer systems at a reasonable time and cost by only employing parallel
global optimization algorithms.
\par
The general state of the art in the field of global optimization has been presented in a
number of key monographs (Törn, A., Žilinskas, A., 1989), (Horst, R., Tuy, H., 1990), (Zhigljavsky, A. A., 1991), (Pintér, J. D., 1996), (Strongin, R. G., Sergeyev, Y. D., 2000), (Locatelli, M., Schoen, F., 2013), (Floudas, C. A., Pardalos, M. P., 2016), etc.
The development of optimization methods that use high-performance computer systems to
solve time-consuming global optimization problems is an area receiving extensive
attention – see, for example, (Censor, Y., Zenios, S. A., 1998), (Strongin, R. G., Sergeyev, Y. D., 2000), (Ciegis, R., Henty, D., Kågström, B., Zilinskas, J., 2009), (Luque, G., Alba, E., 2011), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
\par
The theoretical results obtained provide efficient solutions to many applied global
optimization problems in various fields of scientific and technological applications – see,
for example, (Floudas, C. A., Pardalos, M. P., 1996), (Pintér, J. D., 1996), (Luque, G., Alba, E., 2011), (Fasano, G., Pintér, J. D., (2013), (Locatelli, M., Schoen, F., 2013), (Pardalos, M. P., Zhigljavsky, A. A., Zilinskas, J., 2016), etc.
\par
At the same time, the practical implementation of these algorithms for multiextremal
optimization within the framework of industrial software systems is quite limited.
In many cases, software implementations of global optimization algorithms are experimental in
nature and are just used by the developers themselves to obtain results from computational
experiments required for scientific publication. To a large extent, this situation
originates from high development costs for professional software systems that can be
used by numerous users. In addition, global optimization problems could rarely be
solved in an automatic mode because of their complexity. As a rule, the user should
actively control the global search process which implies an adequate level of
qualification in the field of optimization (particularly, the user should know and
understand global optimization methods well).
\par
On the market for global optimization software, one can select from the following systems:
\begin{itemize}
\item LGO (Lipschitz Global Optimization) is designed to solve global optimization
problems for which the criteria and constraints satisfy the Lipschitz condition (see, for example, (Pintér, J. D., 1996)).
The system is a commercial product based on diagonal extensions of one-dimensional
multiextremal optimization algorithms.
\item GlobSol (Kearfott, R. B., 2009) is oriented towards solving global optimization problems as well as
systems of nonlinear equations. The system includes interval methods based on the branch and
bound method. There are some extensions of the system for parallel computations, and it is available to use for free.
\item LINDO (Lin, Y., Schrage.L., 2009) is features by a wide spectrum of problem solving mehtods which
it can be used for – these include linear, integer, stochastic, nonlinear, and global
optimization problems. The ability to interact with the Microsoft Excel software
environment is a key feature of the system. The system is widely used in practical
applications and is available to use for free.
\item IOSO (Indirect Optimization on the basis of Self-Organization) is oriented
toward solving of a wide class of the extremal problems including global optimization
problems (see, for example, (Egorov, I. N., Kretinin, G. V., Leshchenko, I. A., Kuptzov, S. V., 2002)). The system is widely used to
solve applied problems in various fields. There are versions of the system for parallel
computational systems. The system is a commercial product, but is available for trial use.
\item MATLAB Global Optimization Toolkit (see, for example, (Venkataraman, P., 2009)) includes a wide spectrum
of methods for solving the global optimization problems, including multistart methods,
global pattern search, simulated annealing methods, etc. The library is compatible to the
TOMLAB system (see, for example, (Holmström, K., Edvall, M. M., 2004)), which is an additional extension the widely-used MATLAB.
It is also worth noting that similar libraries for solving global optimization problems are
available for MathCAD, Mathematica, and Maple systems as well.
\item BARON (Branch-And-Reduce Optimization Navigator) is designed to solve continuous
integer programming and global optimization problems using the branch and bound method (Sahinidis, N. V., 1996).
BARON is included in the GAMS (General Algebraic Modeling System) system used extensively – see (Bussieck, M. R., Meeraus, A., 2004).
\item Global Optimization Library in R is a large collection of optimization methods
implemented in the R language (see, for example, (Mullen, K. M., 2014)). Among these methods, are stochastic and deterministic global optimization algorithms,
the branch and bound method, etc.
\end{itemize}
\par
The list provided above is certainly not exhaustive – additional information on software
systems for a wider spectrum of optimization problems can be obtained, for example,
in (Rios, L. M., Sahinidis, N. V., 2013), (Mongeau, M.,  Karsenty, H., Rouzé, V., Hiriart-Urruty, J. B., 2000), (Pintér, J. D., 2009), etc.
Nevertheless, even from such a short list the following conclusions can be drawn (see also Liberti, L., 2006):
\begin{itemize}
\item  the collection of available global optimization software systems for practical use is insufficient,
\item the availability of numerous methods through these systems allows complex
optimization problems to be solved in a number of cases, however, it requires a rather
high level of user knowledge and understanding in the field of global optimization,
\item the use of the parallel computing to increase the efficiency in solving complex
time-consuming problems is limited, therefore, the computational potential of modern
supercomputer systems is very poorly utilized.
\end{itemize}
\par
In this paper, a novel Globalizer software system is considered. The development of the
system was conducted based on the information-statistical theory of multiextremal
optimization aimed at developing efficient parallel algorithms for global search – see, for example, (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A., 2013).
The advantage of the Globalizer is that the system is designed to solve time-consuming
multiextremal optimization problems. In order to obtain global optimized solutions
within a reasonable time and cost, the system efficiently uses modern high-performance computer systems.
\par
This paper is further structured as follows. In Section \ref{sec:problem}, a statement of the multidimensional
global optimization problem is presented, and an approach to reducing these to one-dimensional
optimization problems is described. In Section \ref{sec:parallel}, parallel computation schemes are presented,
and parallel optimization methods are described. In Section \ref{sec:arch}, the Globalizer architecture
is examined. In Section \ref{sec:experiments}, the results are presented from computational experiments
that confirm the system’s high level of efficiency. Finally, Section \ref{sec:concl} presents the
conclusions and some ideas for future research.

\section{Multidimensional Global Optimization Problems and Dimension Reduction}
\label{sec:problem}
In this paper, the core class\footnote{In general, the Globalizer can be applied for solving
multicriterial multiextremal multidimensional optimization problems with nonlinear constraints.} of
optimization problems which can be solved using the
Globalizer is examined. This involves multidimensional global optimization problems
without constraints, which can be defined in the following way:
\begin{equation}
\label{eq:task}
\begin{array}{cr}\\
  \varphi(y^*)=\min\{\varphi(y):y\in D\} \\
  D=\{y\in \mathbf{R}^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}
\end{array}
\end{equation}
where the objective function (y) satisfies the Lipschitz condition
\begin{equation}
\label{eq:lip}
|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,
\end{equation}
where \(L>0\) is the Lipschitz constant, and \(||*||\) denotes the norm in \(\mathbf{R}^N\) space.
\par
Let us further assume that the minimized function \(\varphi(y)\) is defined as a
computational procedure, according to which the value \(\varphi(y)\) can be calculated
for any value of vector \(y\in D\) (let us further call the process of obtaining a
value for the minimized function a trial). As a rule, this procedure is computational-costly i.e.
the overall costs of solving optimization problem (\ref{eq:task}) are determined,
first of all, by the number of trials executed. It is also worth noting the essence of the
assumption on satisfying the Lipschitz condition, since one can construct an estimate
of the global minimum based on a finite number of computed values from the optimized function in this case only.
\par
As has been previously shown by many researchers, finding numerical estimates of globally
optimal extrema implies constructing coverage of the search domain \(D\). As a result,
the computational costs of solving global optimization problems are already very high
even for a small number of varied parameters (the dimensionality of the problem).
A notable reduction in the volume of computations can be achieved when the coverages of the
search domain being obtained are non-uniform, i.e. the series of trial points is dense
only in terms of its nearness to the sought-after globally optimized variants.
The generation of such non-uniform coverages could only be provided in an adaptive
way when the selection of the next trial points is determined by the search information
(the preceding trial points and the values of the minimized function at these points)
obtained in the course of computation. This necessary condition considerably
complicates the computational schemes for global optimization methods, since it
implies a complex analysis of a large amount of multidimensional search information.
As a result, many optimization algorithms use, to some extent, various methods of dimensional reduction.
\par
Within the framework of the information-statistical approach, Peano curves (or evolvents)
\(y(x)\) mapping the interval \([0, 1]\) onto an \(N\)-dimensional hypercube \(D\) are
unambiguously used for dimensional reduction (see, for example, Strongin, R. G., Sergeyev, Y. D., 2000).
\par
As a result of the reduction, the initial multidimensional global optimization
problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
\begin{equation}
\label{eq:oneDimTask}
\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}
\end{equation}
\par
The dimensional reduction scheme considered above superimposes a multidimensional
problem with a minimized Lipschitz function to a one-dimensional problem with a
corresponding minimized function to satisfy the uniform Hölder condition (see Strongin, R. G., 1978, Strongin, R. G., Sergeyev, Y. D., 2000) i.e.
\begin{equation}
\label{eq:holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
\end{equation}
where the constant \(H\) is defined by the relationship \(H=4L\sqrt{N}\), \(L\) is the
Lipschitz constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem (\ref{eq:task}).
\par
The algorithms for the numerical construction of the Peano curve approximations are presented in (Strongin, R. G., Sergeyev, Y. D., 2000).
As an illustration, an approximation of a Peano curve for the third level of density is
shown in Figure \ref{fig:peanoC}. The curve shown in Figure \ref{fig:peanoC} demonstrates the winding order of a
two-dimensional domain; the precision of the Peano curve approximation is determined by the
level of density used in the construction.

\begin{figure}[h!]
    \centering
		\includegraphics[width=0.85\textwidth]{pictures/peanoC.eps}
		\caption{A Peano curve approximation for the third level of}
    \label{fig:peanoC}
\end{figure}

\par
The computational scheme obtained as a result of the dimensional reduction consists of the following (see Figure \ref{fig:peanoCUsage}):
\begin{itemize}
  \item The optimization algorithm performs the minimization of the reduced one-dimensional function \(\varphi(y(x)\),
  \item After determining the next trial point \(x\), a multidimensional image \(у\) in the mapping \(y(x)\) is calculated,
  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the multidimensional point \(у\in D\),
  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
\end{itemize}

\begin{figure}[h!]
    \centering
		\includegraphics[width=0.55\textwidth]{pictures/peanoCUsage.eps}
		\caption{The computational scheme for obtaining the value of the reduced one-dimensional function \(\varphi(y(x)\)}
    \label{fig:peanoCUsage}
\end{figure}

\section{The Parallelization Approach and Global Optimization Algorithms}
\label{sec:parallel}
Let us consider the parallelization methods used widely in the theoretical and
practical application of parallel computing within the context of global optimization problems:
\begin{itemize}
  \item The distribution of the search domain \(D\) among the available computing units
  (data parallelization scheme). In the case of optimization problems, this approach is
  insufficient, since in organizing such computations the subdomain, which contain the
  sought global minimum, will be processed by only one processor, and, therefore,
  all of the remaining processors would perform the excess computations.
  \item The parallelization of the computational schemes for optimization algorithms
  (task parallelization scheme). This approach is also insufficient since the direct
  computational costs for executing optimization algorithms are relatively low
  (the majority of computations in the course of a global search are represented by
  calculations of the optimized function values due to the initial assumption of
  considerable computational costs for such calculations).
  \item The parallelization of the computations executed in order to obtain values
  for the optimized function. This approach can be applied since the most computationally
  costly part of the global search process will be parallelized. However, this method is not
  featured by the generality (the development of parallelization methods is to be performed
  every time from scratch while solving each particular optimization problem).
\end{itemize}
\par
Within the framework of information-statistical theory, a general approach to
parallelization computations when solving global optimization problems has been
proposed --- \textit{the parallelism of computations is provided by means of simultaneously
computing the values of the minimized function \(\varphi(y)\) at several different
points within the search domain \(D\)} – see, for example, (Strongin, R. G., Sergeyev, Y. D., 2000), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
This approach provides parallelizaion for the most costly part of computations in the global search process.
\par
The global optimization algorithms implemented in Globalizer will be described step-by-step below.
In Subsection \ref{subsec:corepar}, the core sequential multidimensional algorithm of global search (MAGS) will be presented.
In Subsection \ref{subsec:sharedpar}, a parallel generalization of the MAGS algorithm for parallel computations on
computer systems with shared memory will be described. In Subsection \ref{subsec:distribpar}, the scheme
for parallel computations by multiprocessor systems with distributed memory will be provided.

\subsection{Core Multidimensional Generalized Algorithm of Global Search}
\label{subsec:corepar}
The information-statistical theory of global search formulated in (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000)
has served as a basis for the development of a large number of efficient multiextremal
optimization methods --- see, for example, (Gergel, V. P., 1996, 1997), (Gergel, V. P., Strongin, R. G., 2003, 2005), (Grishagin, V. A., Strongin, R. G., 1984), (Sergeyev Y. D., 1995, 1999), (Sergeyev, Y. D., Grishagin V. A., 1994, 2001), (Sergeyev, Y. D., Strongin, R. G., Lera, D., 2013), (Barkalov K. A., Gergel V. P., 2014), etc.
\par
Multidimensional Algorithm of Global Search (MAGS) established the basis for the
methods applied in Globalizer. The general computational scheme of MAGS can be
presented as follows --- see (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000).
\par
Let us introduce a simpler notation for the problem being solved on a computing node
\begin{equation}
\label{eq:oneDimTask}
f(x) = \varphi(y(x)):x\in [0;1]\}
\end{equation}
\par
The initial iteration of the algorithm is performed at an arbitrary point \(x^1(0,1)\).
Let us assume further \(k\), \(k>1\), iterations of a global search are to be completed.
The selection of the trial point \(k+1\) for the next iteration is performed according to the following rules.
\subsection{Parallel Computations for Systems with Shared Memory}
\label{subsec:sharedpar}

\subsection{Parallel Computations for Systems with Distributed Memory}
\label{subsec:distribpar}

\section{Globalizer System Architecture}
\label{sec:arch}
\section{Results of Numerical Experiments}
\label{sec:experiments}
\subsection{Parallel Computations Using Intel Xeon Phi processors}
\subsection{Parallel Computations Using General Purpose Graphic Processors}

\section{Conclusions}
\label{sec:concl}
In this paper, the Globalizer global optimization software system was examined for
implementing a general scheme for the parallel solution of globally-optimized decision
making using a wide spectrum of computer systems: with shared and distributed memory,
NVIDIA graphic processors and Intel Xeon Phi coprocessors. This approach enables the
efficient use of modern supercomputer systems to solve the most computational costly
multiextremal optimization problems. The computational experiments have confirmed the
prospects for this proposed approach.

\section*{Acknowledgements}
This research was supported by the Russian Science Foundation, project No 16-11-10150
``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance.''

\end{document}

### 2016-08-04 2016/GOM/paper.tex
Update GOM paper

@@
 \section{Introduction}
+\label{sec:intro}
 Global (or multiextremal) optimization problems are among the most complex problems
@@
 \begin{equation}
-\label{eq:oneDimTask}
-f(x) = \varphi(y(x)):x\in [0;1]\}
+\label{eq:oneDimFunc}
+f(x) = \varphi(y(x)):x\in [0,1]\}
 \end{equation}
@@
 \label{sec:arch}
+As previously mentioned in Section \ref{sec:intro}, the development of software systems
+for global optimization is complex professional work. The optimization systems should
+include advanced optimization methods, provide for the accumulation and efficient
+processing of search information, include various forms of visualization for the results,
+provide tools for interacting with users, etc. All the features pointed out above
+become much more complicated when organizing parallel computations using high-performance
+multiprocessor systems of varying architectures.
+\par
+The Globalizer considered in this paper expands the family of global optimization
+software systems successively developed by the authors during the past several years.
+One of the first developments was the SYMOP multiextremal optimization system (see Gergel, V. P., 1993),
+which has been successfully applied for solving many optimization problems. A special
+place is occupied by the ExaMin system (see, for example, (Barkalov, K., Gergel V., 2015)),
+which was developed and used extensively to investigate the application of novel parallel
+algorithms to solve global optimization problems using high-performance multiprocessor computing systems.
+
+\begin{figure}[h!]
+    \centering
+		\includegraphics[width=0.75\textwidth]{pictures/globalizerScheme.eps}
+		\caption{The Globalizer system architecture (Blocks 1-2, 5-7 have been implemented; Blocks 3-4 and 8-11 are under development)}
+    \label{fig:globalizerScheme}
+\end{figure}
+
+\par
+The Globalizer system architecture is presented in Figure \ref{fig:globalizerScheme}. The structural components are:
+\begin{itemize}
+  \item Block 0 consists of the procedures for computing the function values (criteria
+  and constraints) for the optimization problem being solved; this is an external
+  block with respect to the Globalizer system and is provided by the user per a predefined interface.
+  \item Blocks 1-4 form the optimization subsystem and solve the global optimization
+  problems (Block 1), nonlinear programming (Block 2), multicriterial optimization (Block 3),
+  and general decision making problems (Block 4). It is worth noting the sequential
+  scheme of interaction between these components --- the decision making problems are
+  solved using the multicriterial optimization block, which, in turn, uses the nonlinear programming block, etc.
+  \item Block 5 is a subsystem for accumulating and processing the search information;
+  this is one of the basic subsystems --- the amount of search information for complex
+  optimization problems may appear to be quite large on the one hand, but, on the other
+  hand, the efficiency of the global optimization methods depends to a great extent on how
+  completely all of the available search data is utilized.
+  \item Block 6 contains the dimensional reduction procedures based on Peano evolvents;
+  the optimization blocks solve the reduced (one-dimensional) optimization problems;
+  Block 6 provides interaction between the optimization blocks and the initial
+  multidimensional optimization problem.
+  \item Block 7 organizes the choice of parallel computation schemes in the Globalizer
+  system subject to the computing system architecture employed (the numbers of cores in the
+  processors, the availability of shared and distributed memory, the availability of
+  accelerators for computations, etc.) and the global optimization methods applied.
+  \item Block 8 is responsible for managing the parallel processes when performing the
+  global search (determining the optimal configuration of parallel processes, distributing
+  the processes between computer elements, balancing computational loads, etc.).
+  \item Block 9 is a common managing subsystem, which fully controls the whole
+  computational process when solving global optimization problems.
+  \item Block 10 is responsible for organizing the dialog interaction with users for
+  stating the optimization problem, adjusting system parameters (if necessary), and
+  visualizing and presenting the global search results.
+  \item Block 11 is a set of tools for visualizing and presenting the global search results;
+  the availability of tools for visually presenting the computational results enables
+  the user to provide efficient control over the global optimization process.
+\end{itemize}
+%Figure 3. The Globalizer system architecture (Blocks 1-2, 5-7 have been implemented; Blocks 3-4 and 8-11 are under development)
+\par
+The Globalizer system comes in two different options:
+\begin{itemize}
+  \item Globalizer Pro is a research version of the system providing a full set of
+  capabilities that are oriented toward solving of the most complicated global optimization problems.
+  \item Globalizer Lite is a limited version of the Globalizer system accessible for free for
+  solving global optimization problems with a medium level of difficulty, and is accessible free of charge.
+\end{itemize}
 \section{Results of Numerical Experiments}

### 2016-08-04 2016/GOM/paper.tex
Update GOM paper

--- 2016-08-03 2016/GOM/paper.tex
@@
 \author{
-\name{V. P. Gergel\textsuperscript{a}$^{\ast}$\thanks{$^\ast$Corresponding author. Email: latex.helpdesk@tandf.co.uk}
+\name{V. P. Gergel\textsuperscript{a}$^{\ast}$\thanks{$^\ast$Corresponding author. Email: gergel@unn.ru}
 and K. A. Barkalov\textsuperscript{b} and A. V. Sysoev}
 \affil{\textsuperscript{a}Taylor \& Francis, 4 Park Square, Milton Park, Abingdon, UK;
-\textsuperscript{b}Institut f\"{u}r Informatik, Albert-Ludwigs-Universit\"{a}t, Freiburg, Germany}
-\received{v5.0 released July 2015}
+\textsuperscript{b}State University of Nizhny Novgorod, Nizhny Novgorod, Russia}
 }
@@
 		\includegraphics[width=0.55\textwidth]{pictures/peanoCUsage.eps}
-		\caption{The computational scheme for obtaining the value of the reduced one-dimensional function \(\varphi(y(x)\)}
+		\caption{The computational scheme for obtaining the value of the reduced one-dimensional function \(\varphi(y(x))\)}
     \label{fig:peanoCUsage}
@@
 The selection of the trial point \(k+1\) for the next iteration is performed according to the following rules.
+\par
+\textit{Rule 1}. To renumber the points for the preceding trials by the lower indices in order of increasing value of coordinates
+\begin{equation}
+  \label{step1}
+0=x_0<x_1<\dotsc<x_{k+1}=1
+\end{equation}
+the points \(x_0\), \(x_k\) were introduced additionally for convenience in further explanation,
+the values of the minimized function \(z_0\), \(z_k\) at these points are undefined.
+\par
+\textit{Rule 2}. To compute a current estimate of the Hölder constant \(H\) from (\ref{eq:holder})
+\begin{equation}
+  \label{step2}
+m=\max_{1\leqslant i\leqslant k}\dfrac{|z_i-z_{i-1}|}{\rho_i},
+\begin{matrix}
+    M =
+    \left\{
+    \begin{matrix}
+    r\cdot u,m>0 \\
+    1,m=0
+    \end{matrix} \right.
+    \end{matrix}
+\end{equation}
+as the maximum of the relative differences of the minimized function \(f(x)\) from (\ref{eq:oneDimFunc})
+on the set of executed trial points \(x_i,1\leqslant i\leqslant k\) from (\ref{step1}). Further, \(\rho_i=(x_i-x_{i-1})^\frac{1}{N},1\leqslant i\leqslant k+1\).
+The constant \(r\), \(r>1\), is the \textit{parameter} for the algorithm.
+\par
+\textit{Rule 3.} For each interval \((x_{i-1},x_i),1\leqslant i\leqslant k+1\), compute the characteristics \(R(i)\) where
+\begin{equation}
+\label{step3}
+\begin{array}{cr}\\
+  R(i)=\rho_i+\dfrac{(z_i-z_{i-1})^2}{M^2\rho_i}-2\dfrac{z_i+z_{i-1}}{M},\quad 1<i<k+1 \\
+  R(i)=2\rho_1-4\dfrac{z_1}{M},\quad i=1 \\
+  R(i)=2\rho_{k+1}-4\dfrac{z_k}{M},\quad i=k+1
+\end{array}
+\end{equation}
+\par
+\textit{Rule 4.} To determine the interval with the maximum characteristic
+\begin{equation}
+\label{step4}
+R(t)=\max_{1\leqslant i \leqslant k+1}R(i)
+\end{equation}
+\par
+\textit{Rule 5.} To execute a new trial (computation of the minimized function value \(f(x)\))
+at the point \(x^{k+1}\) located within the interval with the maximum characteristic from (\ref{step4})
+\begin{equation}
+\label{step5}
+\begin{array}{cr}\\
+  x_{k+1}=\dfrac{x_t+x_{t-1}}{2}-sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{\mu}\right]^N,\quad 1<t<k+1 \\
+  x_{k+1}=\dfrac{x_t+x_{t-1}}{2},\quad t=k+1
+\end{array}
+\end{equation}
+\par
+The stop condition by which the trials are terminated, is defined by the inequality
+\begin{equation}
+  \label{eq:stop}
+\rho_t<\varepsilon
+\end{equation}
+for the interval with the maximum characteristic from (\ref{step4}) and \(\varepsilon >0\) is the predefined
+accuracy of the problem solution. If the stop condition is not satisfied, the index \(k\) is incremented by unity,
+and a new global search iteration is executed.
+\par
+In order to explain the algorithm presented above, let us note the following.
+The characteristics \(R(i), 1\leqslant i\leqslant k+1\), calculated according to (\ref{step3})
+could be interpreted as some measures of importance to the intervals with respect to
+location of the global minimum point. Thus, the scheme for selecting the interval
+for executing the next trial described in (\ref{step4}-\ref{step5}) becomes more clear --- the point of every next
+global search iteration is selected within the interval where the global minimum point can most likely be found.
+\par
+The conditions for the algorithm’s convergence described above were examined, for example, in (Strongin, R. G., Sergeyev, Y. D., 2000).
+
 \subsection{Parallel Computations for Systems with Shared Memory}
 \label{subsec:sharedpar}
+Modern supercomputer computational systems consist of plenty of computer nodes,
+which include several multicore processors. At that, random access memory is
+shared for the computing nodes --- the content of any memory element can be read (written)
+by any computer cores available at any arbitrary moment. In most cases, shared memory is
+uniform --- the time characteristics for accessing memory are the same for all computer
+cores and for all memory elements.
+\par
+The following speculations can form the basis for selecting parallel computation
+organization methods. As was mentioned above when describing the MAGS algorithm,
+the characteristics \(R(i),1\leqslant i\leqslant k+1\), calculated according to (\ref{step3})
+can be interpreted as some measures of the importance of intervals with respect to the location of
+the global minimum point. Following this understanding, each successive global search
+iteration is executed within the most important interval with the maximum value of the
+characteristic. So far, in this case it is obvious how to select the other intervals for
+organizing simultaneous computations of the minimized function values at several different
+points within the search domain as well --- these should be the intervals with the next
+magnitudes of characteristics.
+\par
+The computational scheme of the Parallel Multidimensional Algorithm of Global Search for
+computer systems with shared memory (PMAGS) was developed based the approach considered
+above and is practically identical to the MAGS scheme --- the differences consist just in the
+following set of rules --- see (Strongin, R. G., Sergeyev, Y. D., 2000), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
+\par
+\textit{Rule 4'}. To arrange the characteristics of the intervals obtained according to (\ref{step3}) in decreasing order
+\begin{equation}
+\label{step4par}
+R(t_1)\geqslant R(t_2)\geqslant \dots \geqslant R(t_{k})\geqslant R(t_{k+1})
+\end{equation}
+and to select \(p\) intervals with the indices \(t_j,1\leqslant j\leqslant p\), with
+the highest values of characteristics (\(p\) is the number of processors/cores used for the parallel computations).
+\par
+\textit{Rule 5'}. To execute new trials (computations of values for the minimized
+function \(f(x)\)) at the points \(x_{k+j},1\leqslant j\leqslant p\), located in the
+intervals with the highest characteristics from (\ref{step4par})
+\begin{equation}
+\label{step5par}
+\begin{array}{cr}\\
+x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{\mu}\right]^N,\quad 1<t_j<k+1 \\
+x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},\quad t_j=1,t_j=k+1
+\end{array}
+\end{equation}
+\par
+The stop condition for the algorithm (\ref{eq:stop}) should be checked for all
+intervals from (\ref{step4par}), in which the scheduled trials are executed
+\begin{equation}
+  \label{eq:stop}
+\rho_{t_j}<\varepsilon,1\leqslant j\leqslant p
+\end{equation}
+As before, if the stop condition is not fulfilled, the index \(k\) is
+incremented by \(p\), and a new global search iteration is executed.
+\par
+The conditions for the developed parallel algorithm’s convergence and for the
+non-redundancy of parallel computations have been examined in (Strongin, R. G., Sergeyev, Y. D., 2000).
+Thus, in particular, when the conditions for convergence are satisfied, the parallel
+computations are non-redundant as compared to the sequential method when using up to
+\(2N\) processors/cores (where \(N\) is the dimensionality of the global optimization problem being solved).
 
@@
 \label{subsec:distribpar}
+The next level of parallel computations in high-performance computer systems consists of
+using several computational nodes. For that, each computational node has its own separate
+memory and, in this case, the interaction between different nodes can be provided by
+means of data transfer only via the computer communication network.
+\par
+To organize parallel computations for multiprocessor systems with distributed memory, using a set of mappings
+\begin{equation}
+  \label{eq:maps}
+Y_s(x)=\{y^1(x),\dots,y^s(x)\}
+\end{equation}
+instead of applying a single Peano curve \(y(x)\) has been proposed in (Strongin, R. G., 1992, Strongin, R. G., Sergeyev, Y. D., 2000).
+In order to build the set \(Y_s(x)\), several different approaches can be applied.
+Thus, for example, in (Strongin, R. G., 1992) a scheme has been applied by which each
+mapping \(y_i(x)\) from \(Y_s(x)\) is built as the result of some shift along the main
+diagonal of the hypercube \(D\). This way, the set of constructed Peano curves enables the
+close prototypes \(x'\), \(x''\)  to be obtained from the interval \([0,1]\) for any close
+multidimensional images \(y'\), \(y''\) from \(D\) differing in one coordinate,
+only for some mapping \(y_k(x), 1\leqslant k\leqslant s\).
+\par
+Some other methods for constructing multiple mappings have been examined in (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
+\par
+The set of mappings \(Y_s(x)\) from (\ref{eq:maps}), for multidimensional problem (\ref{eq:task})
+generates \(s\) subsidiary information-linked one-dimensional problems of the kind (\ref{eq:oneDimTask}):
+\begin{equation}
+  \label{eq:oneDimProblemSerie}
+  \varphi(y^l(x^l))=\min\{\varphi(y^l(x)):x\in [0,1]\},1\leqslant l\leqslant s
+\end{equation}
+\par
+It is important to note that the family of one-dimensional problems \(\varphi(y^l(x)),1\leqslant l\leqslant s\),
+obtained as a result of the dimensional reduction is information-linked --- the function
+values computed for any problem \(\varphi(y^l(x))\) from the family (\ref{eq:oneDimProblemSerie}) can be used for all
+of the remaining problems of this family.
+\par
+The information compatibility of the problems from the family (\ref{eq:oneDimProblemSerie}) allows the parallel
+computations to be organized in an obvious enough way. Thus, each particular problem can
+be solved by a separate processor in the computing system; the exchange of search information
+between the processors should be performed during the course of the computations.
+As a result of using such a parallelization method, a unified approach for organizing parallel
+computations for a multiprocessor computer systems with distributed, as well as shared
+memory can be developed. This method consists of the following.
+\begin{enumerate}
+  \item The family of reduced one-dimensional information-linked problems (\ref{eq:oneDimProblemSerie}) is
+  distributed among the computational nodes of a multiprocessor system. One problem,
+  as well as several others from within the family, can be allocated to each particular computational node.
+  \item The Parallel Multidimensional Algorithm of Global Search (PMAGS, see Subsection \ref{subsec:sharedpar})
+  is applied to solve the allocated problems from the family (\ref{eq:oneDimProblemSerie})
+  at each computational node, supplemented by the following rules of information interaction.
+  \begin{enumerate}
+    \item Prior to beginning a new trial at any point \(x'\in [0,1]\) for any problem \(\varphi(y_l(x)),1\leqslant l\leqslant s\),
+    the following should be performed:
+    \begin{itemize}
+      \item compute the image \(y'\in D\) for the point \(x'\in [0, 1]\) according to the mapping \(y_l(x)\),
+      \item determine  the prototypes \(x_i',1\leqslant i\leqslant s\), for each
+      problem from the family (\ref{eq:oneDimProblemSerie}),
+      \item send the prototypes \(x_i',1\leqslant i\leqslant s\) to all computational
+      nodes employed in order to exclude the repeated selection of the intervals in
+      which the prototypes fall, for using them to determine the points for new trials.
+      To organize the data transfer, a queue can be formed at each computational
+      node for sending the trial points and receiving the minimized function values at these points.
+    \end{itemize}
+    \item After completing any trial for any problem \(\varphi(y_l(x)),1\leqslant l\leqslant s\), at any point
+    \(x'\in[0,1]\), it is necessary:
+    \begin{itemize}
+      \item to determine all prototypes \(x_i',1\leqslant i\leqslant s\), for the point of the completed trial
+      for each problem from the family (\ref{eq:oneDimProblemSerie}),
+      \item to send the prototypes \(x_i',1\leqslant i\leqslant s\), and the result of the
+      trial \(z'=\varphi(y_l(x'))\) to all computational nodes employed to include the
+      obtained data into the search information processed within the rules of the parallel global search algorithm.
+    \end{itemize}
+    \item Prior to starting the next global search iteration, the algorithm should
+    check the queue of the received data; if there are some data in the queue,
+    they should be included in the search information.
+\end{enumerate}
+\end{enumerate}
+\par
+The possibility of asynchronous data transfer (the computational nodes process received
+data only upon receipt) is a meaningful feature of such a scheme for organizing parallel computations.
+\par
+In addition, there is no single managing node within this scheme. The number of
+computational nodes can vary during the course of the global search, and excluding any
+node does not result in the loss of the sought global minimum of the minimized multiextremal function.
+\par
+This computational scheme generates the Generalized Parallel Multidimensional Algorithm of
+Global Search (GPMAGS) for high-performance computing systems, which may include many
+multiprocessor (multicore) computational nodes with distributed memory as well as
+accelerators for computations based on Intel Xeon Phi processors and on general purpose graphic processors.
+\par
+Additional information on organizing such parallel computation schemes is provided in (Gergel, V. P., Sidorov, S., 2015).
 
@@
 \label{sec:arch}
+As previously mentioned in Section \ref{sec:intro}, the development of software systems
+for global optimization is complex professional work. The optimization systems should
+include advanced optimization methods, provide for the accumulation and efficient
+processing of search information, include various forms of visualization for the results,
+provide tools for interacting with users, etc. All the features pointed out above
+become much more complicated when organizing parallel computations using high-performance
+multiprocessor systems of varying architectures.
+\par
+The Globalizer considered in this paper expands the family of global optimization
+software systems successively developed by the authors during the past several years.
+One of the first developments was the SYMOP multiextremal optimization system (see Gergel, V. P., 1993),
+which has been successfully applied for solving many optimization problems. A special
+place is occupied by the ExaMin system (see, for example, (Barkalov, K., Gergel V., 2015)),
+which was developed and used extensively to investigate the application of novel parallel
+algorithms to solve global optimization problems using high-performance multiprocessor computing systems.
+
+\begin{figure}[h!]
+    \centering
+		\includegraphics[width=0.75\textwidth]{pictures/globalizerScheme.eps}
+		\caption{The Globalizer system architecture (Blocks 1-2, 5-7 have been implemented; Blocks 3-4 and 8-11 are under development)}
+    \label{fig:globalizerScheme}
+\end{figure}
+
+\par
+The Globalizer system architecture is presented in Figure \ref{fig:globalizerScheme}. The structural components are:
+\begin{itemize}
+  \item Block 0 consists of the procedures for computing the function values (criteria
+  and constraints) for the optimization problem being solved; this is an external
+  block with respect to the Globalizer system and is provided by the user per a predefined interface.
+  \item Blocks 1-4 form the optimization subsystem and solve the global optimization
+  problems (Block 1), nonlinear programming (Block 2), multicriterial optimization (Block 3),
+  and general decision making problems (Block 4). It is worth noting the sequential
+  scheme of interaction between these components --- the decision making problems are
+  solved using the multicriterial optimization block, which, in turn, uses the nonlinear programming block, etc.
+  \item Block 5 is a subsystem for accumulating and processing the search information;
+  this is one of the basic subsystems --- the amount of search information for complex
+  optimization problems may appear to be quite large on the one hand, but, on the other
+  hand, the efficiency of the global optimization methods depends to a great extent on how
+  completely all of the available search data is utilized.
+  \item Block 6 contains the dimensional reduction procedures based on Peano evolvents;
+  the optimization blocks solve the reduced (one-dimensional) optimization problems;
+  Block 6 provides interaction between the optimization blocks and the initial
+  multidimensional optimization problem.
+  \item Block 7 organizes the choice of parallel computation schemes in the Globalizer
+  system subject to the computing system architecture employed (the numbers of cores in the
+  processors, the availability of shared and distributed memory, the availability of
+  accelerators for computations, etc.) and the global optimization methods applied.
+  \item Block 8 is responsible for managing the parallel processes when performing the
+  global search (determining the optimal configuration of parallel processes, distributing
+  the processes between computer elements, balancing computational loads, etc.).
+  \item Block 9 is a common managing subsystem, which fully controls the whole
+  computational process when solving global optimization problems.
+  \item Block 10 is responsible for organizing the dialog interaction with users for
+  stating the optimization problem, adjusting system parameters (if necessary), and
+  visualizing and presenting the global search results.
+  \item Block 11 is a set of tools for visualizing and presenting the global search results;
+  the availability of tools for visually presenting the computational results enables
+  the user to provide efficient control over the global optimization process.
+\end{itemize}
+\par
+The Globalizer system comes in two different options:
+\begin{itemize}
+  \item Globalizer Pro is a research version of the system providing a full set of
+  capabilities that are oriented toward solving of the most complicated global optimization problems.
+  \item Globalizer Lite is a limited version of the Globalizer system accessible for free for
+  solving global optimization problems with a medium level of difficulty, and is accessible free of charge.
+\end{itemize}
 \section{Results of Numerical Experiments}
 \label{sec:experiments}
+In order to confirm the operability of the Globalizer system, results from two
+series of computational experiments are presented in this section.
+\par
+The computational experiments were conducted using the Lobachevsky supercomputer at the
+State University of Nizhny Novgorod (operation system --- CentOS 6.4, management system --- SLURM).
+Each supercomputer node had 2 Intel Sandy Bridge E5-2660 processors, 2.2 GHz, with 64 GB RAM.
+Each processor had 8 cores (i.e. a total of 16 CPU cores were available at each node).
+In order to generate the executable program code, the Intel C++ 14.0.2 compiler was used.
+\par
+A number of nodes in the computer system utilized Intel Xeon Phi 5110P processors,
+each containing 60 cores (a total of 240 threads). In addition, the supercomputer also
+included computer nodes equipped with two GPU NVIDIA Kepler K20Х, each providing 14
+stream multiprocessors (a total of 2688 CUDA-cores). In order to generate the executable
+code, the CUDA Toolkit 6.0 was used.
+\par
+The problems generated by the GKLS-generator (Gaviano, M., Lera, D., Kvasov, D. E., Sergeyev, Y. D., 2003)
+were selected for the test problems. This generator allowed for obtaining the multiextremal
+optimization problems with \textit{a priori} known properties: the number of local minima,
+the sizes of their attractors, the global minimum point, the function value at this point, etc.
+In order to simulate the computation costs inherent in the applied optimization problems
+(see, for example, Bastrakov, S., Gergel, V., Meyerov, I., et al, 2013), in all of the experiments
+computing the objective function was complicated by auxiliary computations, not altering the
+form of the function and the position of its minima.
+\par
+We considered the global minimum \(y^*\) in the test problem to be found if the
+algorithm generated a trial point \(y^k\) in the \(\delta\)-nearness of the global
+minimum i. e. \(\Vert y^k-y^* \Vert \leqslant \delta\). The size of the nearness was
+selected as \(\delta=\Vert b-a \Vert\sqrt[N]{\Delta}\) where \(a\) and \(b\) are
+boundaries of the search domain \(D\), \(N\) is the dimensionality of the optimization problem.
+The maximum allowed number of parallel iterations was \(K_{max} = 10000\), and the evolvent
+density parameter was \(m=10\).The computational experiments considered below were first
+carried out using the ExaMin system (see (Barkalov, K., Gergel V., 2015), (Barkalov, K., Gergel, V., Lebedev, I., 2015), (Barkalov, K., Gergel, V., Lebedev, I., Sysoev, A., 2015), (Gergel, V., Lebedev, I., 2015))
+and then were reproduced using the Globalizer system.
+
 \subsection{Parallel Computations Using Intel Xeon Phi processors}
+The results of the numerical experiments with GPMAGS on an Intel Xeon Phi are provided in Table 1.
+The computations were performed using the Simple and Hard function classes with the dimensions equal to 4 and 5.
+The parameters  at  and  at  were used.
+When using the GPMAGS method, the parameter  was selected for the Simple class, and  was selected for the Hard class.
+\par
+In the first series of experiments, serial computations using MAGS were executed.
+The average number of iterations performed by the method for solving a series of problems for each of these classes is shown in row I.
+The symbol “>” reflects the situation where not all problems of a given class were solved by a given method.
+It means that the algorithm was stopped once the maximum allowable number of iterations Kmax was achieved.
+In this case, the Kmax value was used for calculating the average number of iterations corresponding to the lower estimate of this average value.
+The number of unsolved problems is specified in brackets.
+\par
+In the second series of experiments, parallel computations were executed on a CPU.
+The relative “speedup” in iterations achieved is shown in row II; the speedup of parallel computations was measured in relation to the sequential computations (p=1).
+
 \subsection{Parallel Computations Using General Purpose Graphic Processors}

### 2016-08-04 2016/GOM/paper.tex
Update experiments section of GOM paper

@@
 \newtheorem{definition}{Definition}
+\DeclareMathOperator{\sign}{sign}
 
@@
 \begin{array}{cr}\\
-  x_{k+1}=\dfrac{x_t+x_{t-1}}{2}-sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{\mu}\right]^N,\quad 1<t<k+1 \\
+  x_{k+1}=\dfrac{x_t+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{\mu}\right]^N,\quad 1<t<k+1 \\
   x_{k+1}=\dfrac{x_t+x_{t-1}}{2},\quad t=k+1
@@
 \begin{array}{cr}\\
-x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{\mu}\right]^N,\quad 1<t_j<k+1 \\
+x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-\sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{\mu}\right]^N,\quad 1<t_j<k+1 \\
 x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},\quad t_j=1,t_j=k+1
@@
 The results of the numerical experiments with GPMAGS on an Intel Xeon Phi are provided in Table 1.
-The computations were performed using the Simple and Hard function classes with the dimensions equal to 4 and 5.
-The parameters  at  and  at  were used.
-When using the GPMAGS method, the parameter  was selected for the Simple class, and  was selected for the Hard class.
+The computations were performed using the Simple and Hard function classes with the
+dimensions equal to 4 and 5. The parameters \(\Delta=10^{-6}\) at \(N=4\) and \(\Delta=10^{-6}\) at \(N=5\)
+were used. When using the GPMAGS method, the parameter \(r=4.5\) was selected for the
+Simple class, and \(r=5.6\) was selected for the Hard class.
 \par
 In the first series of experiments, serial computations using MAGS were executed.
-The average number of iterations performed by the method for solving a series of problems for each of these classes is shown in row I.
-The symbol “>” reflects the situation where not all problems of a given class were solved by a given method.
-It means that the algorithm was stopped once the maximum allowable number of iterations Kmax was achieved.
-In this case, the Kmax value was used for calculating the average number of iterations corresponding to the lower estimate of this average value.
-The number of unsolved problems is specified in brackets.
+The average number of iterations performed by the method for solving a series of
+problems for each of these classes is shown in row I. The symbol ``\(>\)'' reflects the
+situation where not all problems of a given class were solved by a given method.
+It means that the algorithm was stopped once the maximum allowable number of
+iterations \(K_{max}\) was achieved. In this case, the \(K_{max}\) value was used for
+calculating the average number of iterations corresponding to the lower estimate of
+this average value. The number of unsolved problems is specified in brackets.
 \par
 In the second series of experiments, parallel computations were executed on a CPU.
-The relative “speedup” in iterations achieved is shown in row II; the speedup of parallel computations was measured in relation to the sequential computations (p=1).
+The relative ``speedup'' in iterations achieved is shown in row II; the speedup of
+parallel computations was measured in relation to the sequential computations (\(p=1\)).
+
+\par
+The final series of experiments was executed using a Xeon Phi. The results of these
+computations are shown in row III; in this case, the speedup factor is calculated in
+relation to the PMAGS results on a CPU using eight cores (\(p=8\)). It can be noted that
+relatively high speedup was achieved. For example, solving a problem from the 5-Hard class
+required on average only 633 parallel iterations on the Xeon Phi, whereas the number of
+iterations was more than 37 thousand when using all of the CPU’s computing cores.
 
 \subsection{Parallel Computations Using General Purpose Graphic Processors}
+In the first series of experiments, only one graphic accelerator was used for solving
+100 6-diminsional problems. The method’s reliability parameter \(r=4.5\) was selected.
+\par
+The average problem solving time using a GPU was 10.78 sec. whereas the averaged problem
+solving time using all of the node’s 16 CPU cores was 53.8 sec.; an almost 5-fold improvement was observed.
+\par
+A larger scale computational experiment was carried out for the problems with
+dimensions of \(N=8\) and \(N=10\): a total of 12 nodes within a cluster with 36
+graphic accelerators (three accelerators per node) were employed, therefore, a total
+of 96768 CUDA-cores were employed.
+\par
+The average time to solve the 8-dimensional problem was 405.6 seconds, 5.9 times
+faster than the CPU version of the algorithm. The average time to solve the
+10-dimensional problem was 2,055.8 sec. The CPU was not used to solve the 10-dimensional
+problem because of the high degree of complexity in the computations.
 

### 2016-08-06 2016/GOM/paper.tex
Finish all sections of GOM except references

@@
 \usepackage{subfigure}% Support for small, `sub' figures and tables
+\usepackage{multirow}
+\usepackage{array}
+\usepackage[utf8]{inputenc}
 
@@
 \DeclareMathOperator{\sign}{sign}
+\newcolumntype{K}[1]{>{\centering\arraybackslash}p{#1}}
 
@@
 \end{equation}
-where the objective function (y) satisfies the Lipschitz condition
+where the objective function \(\varphi(y)\) satisfies the Lipschitz condition
 \begin{equation}
@@
 \end{equation}
-where \(L>0\) is the Lipschitz constant, and \(||*||\) denotes the norm in \(\mathbf{R}^N\) space.
+where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbf{R}^N\) space.
 \par
@@
 for any value of vector \(y\in D\) (let us further call the process of obtaining a
-value for the minimized function a trial). As a rule, this procedure is computational-costly i.e.
+value for the minimized function \textit{a trial}). As a rule, this procedure is computational-costly i.e.
 the overall costs of solving optimization problem (\ref{eq:task}) are determined,
@@
 \par
-Within the framework of the information-statistical approach, Peano curves (or evolvents)
+Within the framework of the information-statistical approach, Peano \textit{curves} (or \textit{evolvents})
 \(y(x)\) mapping the interval \([0, 1]\) onto an \(N\)-dimensional hypercube \(D\) are
@@
   \item The optimization algorithm performs the minimization of the reduced one-dimensional function \(\varphi(y(x)\),
-  \item After determining the next trial point \(x\), a multidimensional image \(у\) in the mapping \(y(x)\) is calculated,
-  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the multidimensional point \(у\in D\),
+  \item After determining the next trial point \(x\), a multidimensional image \(y\) in the mapping \(y(x)\) is calculated,
+  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the multidimensional point \(y\in D\),
   \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
@@
     \begin{matrix}
-    r\cdot u,m>0 \\
+    r\cdot m,m>0 \\
     1,m=0
@@
 \par
-\textit{Rule 3.} For each interval \((x_{i-1},x_i),1\leqslant i\leqslant k+1\), compute the characteristics \(R(i)\) where
+\textit{Rule 3}. For each interval \((x_{i-1},x_i),1\leqslant i\leqslant k+1\), compute the characteristics \(R(i)\) where
 \begin{equation}
@@
   R(i)=\rho_i+\dfrac{(z_i-z_{i-1})^2}{M^2\rho_i}-2\dfrac{z_i+z_{i-1}}{M},\quad 1<i<k+1 \\
-  R(i)=2\rho_1-4\dfrac{z_1}{M},\quad i=1 \\
-  R(i)=2\rho_{k+1}-4\dfrac{z_k}{M},\quad i=k+1
+  R(i)=2\rho_i-4\dfrac{z_i}{M},\quad i=1 \\
+  R(i)=2\rho_{i}-4\dfrac{z_{i-1}}{M},\quad i=k+1
 \end{array}
@@
 \par
-\textit{Rule 4.} To determine the interval with the maximum characteristic
+\textit{Rule 4}. To determine the interval with the maximum characteristic
 \begin{equation}
@@
 \par
-\textit{Rule 5.} To execute a new trial (computation of the minimized function value \(f(x)\))
+\textit{Rule 5}. To execute a new trial (computation of the minimized function value \(f(x)\))
 at the point \(x^{k+1}\) located within the interval with the maximum characteristic from (\ref{step4})
@@
 \begin{array}{cr}\\
-  x_{k+1}=\dfrac{x_t+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{\mu}\right]^N,\quad 1<t<k+1 \\
+  x_{k+1}=\dfrac{x_t+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{M}\right]^N,\quad 1<t<k+1 \\
   x_{k+1}=\dfrac{x_t+x_{t-1}}{2},\quad t=k+1
@@
 \begin{array}{cr}\\
-x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-\sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{\mu}\right]^N,\quad 1<t_j<k+1 \\
+x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-\sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\quad 1<t_j<k+1 \\
 x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},\quad t_j=1,t_j=k+1
@@
   \label{eq:oneDimProblemSerie}
-  \varphi(y^l(x^l))=\min\{\varphi(y^l(x)):x\in [0,1]\},1\leqslant l\leqslant s
+  \varphi(y^l(x^*))=\min\{\varphi(y^l(x)):x\in [0,1]\},1\leqslant l\leqslant s
 \end{equation}
@@
 each containing 60 cores (a total of 240 threads). In addition, the supercomputer also
-included computer nodes equipped with two GPU NVIDIA Kepler K20Х, each providing 14
+included computer nodes equipped with two GPU NVIDIA Kepler K20X, each providing 14
 stream multiprocessors (a total of 2688 CUDA-cores). In order to generate the executable
@@
 \subsection{Parallel Computations Using Intel Xeon Phi processors}
-The results of the numerical experiments with GPMAGS on an Intel Xeon Phi are provided in Table 1.
+The results of the numerical experiments with GPMAGS on an Intel Xeon Phi are provided in Table \ref{tab:exResults}.
 The computations were performed using the Simple and Hard function classes with the
@@
 parallel computations was measured in relation to the sequential computations (\(p=1\)).
+
+\begin{table}[h]
+  \centering
+  \label{tab:exResults}
+  \caption{Results of numerical experiments}
+  \begin{tabular}{|l|K{0.3\textwidth}|c|c|c|c|c|c|}
+\hline
+\multicolumn{2}{|c|}{} & \(p\) & \multicolumn{2}{c|}{\(N=4\)} & & \multicolumn{2}{c|}{\(N=5\)} \\
+ \cline{4-5} \cline{7-8}
+ \multicolumn{2}{|c|}{} &  & \multicolumn{2}{c|}{Simple} & & \multicolumn{2}{c|}{Hard} \\
+\hline
+I   & \textbf{Serial computations.} \textit{Average number of iterations} & 1 & 11953 & 25263 & & 15920 & \(>\)148342 (4) \\ \hline
+II  & \textbf{Parallel computations}  \multirow{3}{*}{}
+  & 2 & 2.51 & 2.26 & & 1.19 & 1.36 \\ \cline{3-5} \cline{7-8}
+& of CPU. & 4 & 5.04 & 4.23 & & 3.06 & 2.86 \\ \cline{3-5} \cline{7-8}
+& \textit{Speedup} & 8 & 8.58 & 8.79 & & 4.22 & 6.56 \\
+ \hline
+III & \textbf{Parallel computations} \multirow{3}{*}{}
+  & 60  & 8.13 & 7.32 & & 9.87 & 6.55  \\ \cline{3-5} \cline{7-8}
+&  of Xeon Phi. & 120 & 16.33 & 15.82 & & 15.15 & 17.31 \\ \cline{3-5} \cline{7-8}
+& \textit{Speedup} & 240 & 33.07 & 27.79 & & 38.80 & 59.31 \\
+\hline
+\end{tabular}
+\end{table}
 

### 2016-08-07 2016/GOM/paper.tex
Add references section of GOM paper (WIP)

@@
 The general state of the art in the field of global optimization has been presented in a
-number of key monographs (Törn, A., Žilinskas, A., 1989), (Horst, R., Tuy, H., 1990), (Zhigljavsky, A.
-A., 1991), (Pintér, J.
-D., 1996), (Strongin, R.
-G., Sergeyev, Y.
-D., 2000), (Locatelli, M., Schoen, F., 2013), (Floudas, C.
-A., Pardalos, M.
-P., 2016), etc.
+number of key monographs \cite{zilinskTornGO}, \cite{horstTuyGO}, \cite{zhigljavskyRandGO},
+\cite{pinterGO}, \cite{strSergGO}, \cite{locatelliSchoenGO}, \cite{floudasPardGO}, etc.
 The development of optimization methods that use high-performance computer systems to
@@
 attention – see, for example, (Censor, Y., Zenios, S.
-A., 1998), (Strongin, R.
-G., Sergeyev, Y.
-D., 2000), (Ciegis, R., Henty, D., Kågström, B., Zilinskas, J., 2009), (Luque, G., Alba, E., 2011), (Strongin, R.
+A., 1998), \cite{strSergGO}, (Ciegis, R., Henty, D., Kågström, B., Zilinskas, J., 2009), (Luque, G., Alba, E., 2011), (Strongin, R.
 G., K.
@@
 optimization aimed at developing efficient parallel algorithms for global search – see, for example, (Strongin, R.
-G., 1978), (Strongin, R.
-G., Sergeyev, Y.
-D., 2000), (Strongin, R.
+G., 1978), \cite{strSergGO}, (Strongin, R.
 G., K.
@@
 \(y(x)\) mapping the interval \([0, 1]\) onto an \(N\)-dimensional hypercube \(D\) are
-unambiguously used for dimensional reduction (see, for example, Strongin, R. G., Sergeyev, Y. D., 2000).
+unambiguously used for dimensional reduction (see, for example, \cite{strSergGO}).
 \par
@@
 problem with a minimized Lipschitz function to a one-dimensional problem with a
-corresponding minimized function to satisfy the uniform Hölder condition (see Strongin, R. G., 1978, Strongin, R. G., Sergeyev, Y. D., 2000) i.e.
+corresponding minimized function to satisfy the uniform Hölder condition (see Strongin, R. G., 1978, \cite{strSergGO}) i.e.
 \begin{equation}
@@
 \par
-The algorithms for the numerical construction of the Peano curve approximations are presented in (Strongin, R. G., Sergeyev, Y. D., 2000).
+The algorithms for the numerical construction of the Peano curve approximations are presented in \cite{strSergGO}.
 As an illustration, an approximation of a Peano curve for the third level of density is
@@
 computing the values of the minimized function \(\varphi(y)\) at several different
-points within the search domain \(D\)} – see, for example, (Strongin, R. G., Sergeyev, Y. D., 2000), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
+points within the search domain \(D\)} – see, for example, \cite{strSergGO}, (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
 This approach provides parallelizaion for the most costly part of computations in the global search process.
@@
 \label{subsec:corepar}
-The information-statistical theory of global search formulated in (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000)
+The information-statistical theory of global search formulated in (Strongin, R. G., 1978), \cite{strSergGO}
 has served as a basis for the development of a large number of efficient multiextremal
@@
 methods applied in Globalizer. The general computational scheme of MAGS can be
-presented as follows --- see (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000).
+presented as follows --- see (Strongin, R. G., 1978), \cite{strSergGO}.
 \par
@@
 \par
-The conditions for the algorithm’s convergence described above were examined, for example, in (Strongin, R. G., Sergeyev, Y. D., 2000).
+The conditions for the algorithm’s convergence described above were examined, for example, in \cite{strSergGO}.
 
@@
 above and is practically identical to the MAGS scheme --- the differences consist just in the
-following set of rules --- see (Strongin, R. G., Sergeyev, Y. D., 2000), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
+following set of rules --- see \cite{strSergGO}, (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
 \par
@@
 The conditions for the developed parallel algorithm’s convergence and for the
-non-redundancy of parallel computations have been examined in (Strongin, R. G., Sergeyev, Y. D., 2000).
+non-redundancy of parallel computations have been examined in \cite{strSergGO}.
 Thus, in particular, when the conditions for convergence are satisfied, the parallel
@@
 \end{equation}
-instead of applying a single Peano curve \(y(x)\) has been proposed in (Strongin, R. G., 1992, Strongin, R. G., Sergeyev, Y. D., 2000).
+instead of applying a single Peano curve \(y(x)\) has been proposed in (Strongin, R. G., 1992, \cite{strSergGO}).
 In order to build the set \(Y_s(x)\), several different approaches can be applied.
@@
 
+\bibliographystyle{gOMS}
+\bibliography{paper}
+
 \end{document}

### 2016-08-08 2016/GOM/paper.tex
Add Makefile for GOM paper

@@
 solve time-consuming global optimization problems is an area receiving extensive
-attention – see, for example, (Censor, Y., Zenios, S.
-A., 1998), \cite{strSergGO}, (Ciegis, R., Henty, D., Kågström, B., Zilinskas, J., 2009), (Luque, G., Alba, E., 2011), (Strongin, R.
-G., K.
-A., Gergel, V.
-P., Grishagin, V.
-A., Barkalov, K.
-A.,  2013).
+attention – see, for example, \cite{censorZeniosParGO}, \cite{strSergGO},
+\cite{ciegisHentyParGO}, \cite{luqueAlbaGA}, \cite{stronginGergelBarkalovParGO}.
 \par
@@
 optimization problems in various fields of scientific and technological applications – see,
-for example, (Floudas, C.
-A., Pardalos, M.
-P., 1996), (Pintér, J.
-D., 1996), (Luque, G., Alba, E., 2011), (Fasano, G., Pintér, J.
-D., (2013), (Locatelli, M., Schoen, F., 2013), (Pardalos, M.
-P., Zhigljavsky, A.
-A., Zilinskas, J., 2016), etc.
+for example, \cite{floudasPardalosGOState}, \cite{pinterGO}, \cite{luqueAlbaGA},
+\cite{fasanoPinter2013}, \cite{locatelliSchoenGO}, \cite{pardalosZhigljavskyZilinskas2016}, etc.
 \par
@@
 \item LGO (Lipschitz Global Optimization) is designed to solve global optimization
-problems for which the criteria and constraints satisfy the Lipschitz condition (see, for example, (Pintér, J. D., 1996)).
+problems for which the criteria and constraints satisfy the Lipschitz condition (see, for example, \cite{pinterGO}).
 The system is a commercial product based on diagonal extensions of one-dimensional
@@
 system was conducted based on the information-statistical theory of multiextremal
-optimization aimed at developing efficient parallel algorithms for global search – see, for example, (Strongin, R.
-G., 1978), \cite{strSergGO}, (Strongin, R.
-G., K.
-A., Gergel, V.
-P., Grishagin, V.
-A., Barkalov, K.
-A., 2013).
+optimization aimed at developing efficient parallel algorithms for global search – see, for example, (Strongin, R. G., 1978), \cite{strSergGO}, (\cite{stronginGergelBarkalovParGO}).
 The advantage of the Globalizer is that the system is designed to solve time-consuming
@@
 computing the values of the minimized function \(\varphi(y)\) at several different
-points within the search domain \(D\)} – see, for example, \cite{strSergGO}, (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
+points within the search domain \(D\)} – see, for example, \cite{strSergGO}, \cite{stronginGergelBarkalovParGO}.
 This approach provides parallelizaion for the most costly part of computations in the global search process.
@@
 above and is practically identical to the MAGS scheme --- the differences consist just in the
-following set of rules --- see \cite{strSergGO}, (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
+following set of rules --- see \cite{strSergGO}, \cite{stronginGergelBarkalovParGO}.
 \par
@@
 computations are non-redundant as compared to the sequential method when using up to
-\(2N\) processors/cores (where \(N\) is the dimensionality of the global optimization problem being solved).
+\(2^N\) processors/cores (where \(N\) is the dimensionality of the global optimization problem being solved).
 
@@
 \par
-Some other methods for constructing multiple mappings have been examined in (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
+Some other methods for constructing multiple mappings have been examined in \cite{stronginGergelBarkalovParGO}.
 \par

### 2016-08-09 2016/GOM/paper.tex
Update references in GOM paper

@@
 multiextremal optimization algorithms.
-\item GlobSol (Kearfott, R. B., 2009) is oriented towards solving global optimization problems as well as
+\item GlobSol \cite{kearfott2009} is oriented towards solving global optimization problems as well as
 systems of nonlinear equations. The system includes interval methods based on the branch and
 bound method. There are some extensions of the system for parallel computations, and it is available to use for free.
-\item LINDO (Lin, Y., Schrage.L., 2009) is features by a wide spectrum of problem solving mehtods which
-it can be used for – these include linear, integer, stochastic, nonlinear, and global
+\item LINDO \cite{linSchrage2009} is features by a wide spectrum of problem solving mehtods which
+it can be used for --- these include linear, integer, stochastic, nonlinear, and global
 optimization problems. The ability to interact with the Microsoft Excel software
@@
 toward solving of a wide class of the extremal problems including global optimization
-problems (see, for example, (Egorov, I. N., Kretinin, G. V., Leshchenko, I. A., Kuptzov, S. V., 2002)). The system is widely used to
+problems (see, for example, \cite{iosoDescription}). The system is widely used to
 solve applied problems in various fields. There are versions of the system for parallel
 computational systems. The system is a commercial product, but is available for trial use.
-\item MATLAB Global Optimization Toolkit (see, for example, (Venkataraman, P., 2009)) includes a wide spectrum
+\item MATLAB Global Optimization Toolkit (see, for example, \cite{venkataraman2009}) includes a wide spectrum
 of methods for solving the global optimization problems, including multistart methods,
 global pattern search, simulated annealing methods, etc. The library is compatible to the
-TOMLAB system (see, for example, (Holmström, K., Edvall, M. M., 2004)), which is an additional extension the widely-used MATLAB.
+TOMLAB system (see, for example, \cite{holmstromEdvall2004}), which is an additional extension the widely-used MATLAB.
 It is also worth noting that similar libraries for solving global optimization problems are
@@
 \item BARON (Branch-And-Reduce Optimization Navigator) is designed to solve continuous
-integer programming and global optimization problems using the branch and bound method (Sahinidis, N. V., 1996).
-BARON is included in the GAMS (General Algebraic Modeling System) system used extensively – see (Bussieck, M. R., Meeraus, A., 2004).
+integer programming and global optimization problems using the branch and bound method \cite{sahinidis1996}.
+BARON is included in the GAMS (General Algebraic Modeling System) system used extensively – see \cite{bussieckMeeraus2004}.
 \item Global Optimization Library in R is a large collection of optimization methods
-implemented in the R language (see, for example, (Mullen, K. M., 2014)). Among these methods, are stochastic and deterministic global optimization algorithms,
+implemented in the R language (see, for example, \cite{mullen2014}). Among these methods, are stochastic and deterministic global optimization algorithms,
 the branch and bound method, etc.
@@
 systems for a wider spectrum of optimization problems can be obtained, for example,
-in (Rios, L. M., Sahinidis, N. V., 2013), (Mongeau, M.,  Karsenty, H., Rouzé, V., Hiriart-Urruty, J. B., 2000), (Pintér, J. D., 2009), etc.
-Nevertheless, even from such a short list the following conclusions can be drawn (see also Liberti, L., 2006):
+in \cite{riosSahinidis2013}, \cite{mongeauKarsentyRouze2000}, \cite{pinter2009}, etc.
+Nevertheless, even from such a short list the following conclusions can be drawn (see also \cite{liberti2009}):
 \begin{itemize}
@@
 system was conducted based on the information-statistical theory of multiextremal
-optimization aimed at developing efficient parallel algorithms for global search – see, for example, (Strongin, R. G., 1978), \cite{strSergGO}, (\cite{stronginGergelBarkalovParGO}).
+optimization aimed at developing efficient parallel algorithms for global search --- see, for example, \cite{strongin1978}, \cite{strSergGO}, (\cite{stronginGergelBarkalovParGO}).
 The advantage of the Globalizer is that the system is designed to solve time-consuming
@@
 problem with a minimized Lipschitz function to a one-dimensional problem with a
-corresponding minimized function to satisfy the uniform Hölder condition (see Strongin, R. G., 1978, \cite{strSergGO}) i.e.
+corresponding minimized function to satisfy the uniform Hölder condition (see \cite{strongin1978}, \cite{strSergGO}) i.e.
 \begin{equation}
@@
 \label{subsec:corepar}
-The information-statistical theory of global search formulated in (Strongin, R. G., 1978), \cite{strSergGO}
+The information-statistical theory of global search formulated in \cite{strongin1978}, \cite{strSergGO}
 has served as a basis for the development of a large number of efficient multiextremal
@@
 methods applied in Globalizer. The general computational scheme of MAGS can be
-presented as follows --- see (Strongin, R. G., 1978), \cite{strSergGO}.
+presented as follows --- see \cite{strongin1978}, \cite{strSergGO}.
 \par

### 2016-08-10 2016/GOM/paper.tex
Update references in GOM paper

@@
 has served as a basis for the development of a large number of efficient multiextremal
-optimization methods --- see, for example, (Gergel, V.
-P., 1996, 1997), (Gergel, V.
-P., Strongin, R.
-G., 2003, 2005), (Grishagin, V.
-A., Strongin, R.
-G., 1984), (Sergeyev Y.
-D., 1995, 1999), (Sergeyev, Y.
-D., Grishagin V.
-A., 1994, 2001), (Sergeyev, Y.
-D., Strongin, R.
-G., Lera, D., 2013), (Barkalov K.
-A., Gergel V.
-P., 2014), etc.
+optimization methods --- see, for example, \cite{gergel1996}, \cite{gergel1997},
+\cite{gergelStrongin2003}, \cite{gergelStrongin2005}, \cite{grishaginStrongin1984},
+\cite{sergeyev1995}, \cite{sergeyev1999}, \cite{sergeyevGrishagin1994}, \cite{sergeyevGrishagin2001}
+\cite{sergeyevStronginLera2013}, \cite{barkalovGergel2014}, etc.
 \par
@@
 \end{equation}
-instead of applying a single Peano curve \(y(x)\) has been proposed in (Strongin, R. G., 1992, \cite{strSergGO}).
+instead of applying a single Peano curve \(y(x)\) has been proposed in \cite{strongin1992}, \cite{strSergGO}.
 In order to build the set \(Y_s(x)\), several different approaches can be applied.
-Thus, for example, in (Strongin, R. G., 1992) a scheme has been applied by which each
+Thus, for example, in \cite{strongin1992} a scheme has been applied by which each
 mapping \(y_i(x)\) from \(Y_s(x)\) is built as the result of some shift along the main
@@
 \par
-Additional information on organizing such parallel computation schemes is provided in (Gergel, V. P., Sidorov, S., 2015).
+Additional information on organizing such parallel computation schemes is provided in \cite{gergelSidorov2015}.
 
@@
 software systems successively developed by the authors during the past several years.
-One of the first developments was the SYMOP multiextremal optimization system (see Gergel, V. P., 1993),
+One of the first developments was the SYMOP multiextremal optimization system \cite{gergel1993},
 which has been successfully applied for solving many optimization problems. A special
-place is occupied by the ExaMin system (see, for example, (Barkalov, K., Gergel V., 2015)),
+place is occupied by the ExaMin system (see, for example, \cite{barkalovGergel2015}),
 which was developed and used extensively to investigate the application of novel parallel
@@
 \end{itemize}
+
 \section{Results of Numerical Experiments}
@@
 \par
-The problems generated by the GKLS-generator (Gaviano, M., Lera, D., Kvasov, D. E., Sergeyev, Y. D., 2003)
+The problems generated by the GKLS-generator \cite{gavianoKvasovLeraSergeev2003}
 were selected for the test problems. This generator allowed for obtaining the multiextremal
@@
 The maximum allowed number of parallel iterations was \(K_{max} = 10000\), and the evolvent
-density parameter was \(m=10\).The computational experiments considered below were first
-carried out using the ExaMin system (see (Barkalov, K., Gergel V., 2015), (Barkalov, K., Gergel, V., Lebedev, I., 2015), (Barkalov, K., Gergel, V., Lebedev, I., Sysoev, A., 2015), (Gergel, V., Lebedev, I., 2015))
+density parameter was \(m=10\). The computational experiments considered below were first
+carried out using the ExaMin system (see \cite{barkalovGergel2015}, \cite{barkalovGergelLebedev2015},
+\cite{barkalovGergelLebedevSysoev2015}, \cite{gergelLebedev2015})
 and then were reproduced using the Globalizer system.

### 2016-08-16 2016/GOM/paper.tex
Fix authors list in GOM paper
Some minor corrections

@@
 \author{
-\name{V. P. Gergel\textsuperscript{a}$^{\ast}$\thanks{$^\ast$Corresponding author. Email: gergel@unn.ru}
-and K. A. Barkalov\textsuperscript{b} and A. V. Sysoev}
-\affil{\textsuperscript{a}Taylor \& Francis, 4 Park Square, Milton Park, Abingdon, UK;
-\textsuperscript{b}State University of Nizhny Novgorod, Nizhny Novgorod, Russia}
+\name{V.P. Gergel and K.A. Barkalov$^{\ast}$\thanks{$^\ast$Corresponding author. Email: barkalov@vmk.unn.ru} and A.V. Sysoyev}
+\affil{State University of Nizhny Novgorod, Nizhny Novgorod, Russia}
 }
@@
   \varphi(y^*)=\min\{\varphi(y):y\in D\} \\
-  D=\{y\in \mathbf{R}^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}
+  D=\{y\in \mathbf{R}^N:a_i\leqslant y_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}
 \end{array}
@@
 problem with a minimized Lipschitz function to a one-dimensional problem with a
-corresponding minimized function to satisfy the uniform Hölder condition (see \cite{strongin1978}, \cite{strSergGO}) i.e.
+corresponding minimized function to satisfy the uniform H{\"o}lder condition (see \cite{strongin1978}, \cite{strSergGO}) i.e.
 \begin{equation}
@@
 		\includegraphics[width=0.85\textwidth]{pictures/peanoC.eps}
-		\caption{A Peano curve approximation for the third level of}
+		\caption{A Peano curve approximation for the third level}
     \label{fig:peanoC}
@@
 \begin{itemize}
-  \item The optimization algorithm performs the minimization of the reduced one-dimensional function \(\varphi(y(x)\),
+  \item The optimization algorithm performs the minimization of the reduced one-dimensional function \(\varphi(y(x))\),
   \item After determining the next trial point \(x\), a multidimensional image \(y\) in the mapping \(y(x)\) is calculated,
@@
 \label{eq:oneDimFunc}
-f(x) = \varphi(y(x)):x\in [0,1]\}
+f(x) = \varphi(y(x)):x\in [0,1].
 \end{equation}
@@
   \label{step1}
-0=x_0<x_1<\dotsc<x_{k+1}=1
+0=x_0<x_1<\dotsc<x_{k+1}=1.
 \end{equation}
-the points \(x_0\), \(x_k\) were introduced additionally for convenience in further explanation,
+The points \(x_0\), \(x_k\) were introduced additionally for convenience in further explanation,
 the values of the minimized function \(z_0\), \(z_k\) at these points are undefined.
 \par
-\textit{Rule 2}. To compute a current estimate of the Hölder constant \(H\) from (\ref{eq:holder})
+\textit{Rule 2}. To compute a current estimate of the H{\"o}lder constant \(H\) from (\ref{eq:holder})
 \begin{equation}
@@
 \begin{array}{cr}\\
-  R(i)=\rho_i+\dfrac{(z_i-z_{i-1})^2}{M^2\rho_i}-2\dfrac{z_i+z_{i-1}}{M},\quad 1<i<k+1 \\
-  R(i)=2\rho_i-4\dfrac{z_i}{M},\quad i=1 \\
-  R(i)=2\rho_{i}-4\dfrac{z_{i-1}}{M},\quad i=k+1
+  R(i)=\rho_i+\dfrac{(z_i-z_{i-1})^2}{M^2\rho_i}-2\dfrac{z_i+z_{i-1}}{M},\quad 1<i<k+1, \\
+  R(i)=2\rho_i-4\dfrac{z_i}{M},\quad i=1, \\
+  R(i)=2\rho_{i}-4\dfrac{z_{i-1}}{M},\quad i=k+1.
 \end{array}
@@
 \label{step4}
-R(t)=\max_{1\leqslant i \leqslant k+1}R(i)
+R(t)=\max_{1\leqslant i \leqslant k+1}R(i).
 \end{equation}
@@
 \begin{array}{cr}\\
-  x_{k+1}=\dfrac{x_t+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{M}\right]^N,\quad 1<t<k+1 \\
-  x_{k+1}=\dfrac{x_t+x_{t-1}}{2},\quad t=k+1
+  x_{k+1}=\dfrac{x_t+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{M}\right]^N,\quad 1<t<k+1, \\
+  x_{k+1}=\dfrac{x_t+x_{t-1}}{2},\quad t=k+1.
 \end{array}
@@
 \begin{array}{cr}\\
-x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-\sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\quad 1<t_j<k+1 \\
-x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},\quad t_j=1,t_j=k+1
+x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-\sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\quad 1<t_j<k+1, \\
+x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},\quad t_j=1,t_j=k+1.
 \end{array}
@@
   \label{eq:stop}
-\rho_{t_j}<\varepsilon,1\leqslant j\leqslant p
+\rho_{t_j}<\varepsilon,1\leqslant j\leqslant p.
 \end{equation}
@@
   \label{eq:oneDimProblemSerie}
-  \varphi(y^l(x^*))=\min\{\varphi(y^l(x)):x\in [0,1]\},1\leqslant l\leqslant s
+  \varphi(y^l(x^*))=\min\{\varphi(y^l(x)):x\in [0,1]\},1\leqslant l\leqslant s.
 \end{equation}

### 2016-09-18 2016/YSC/ysc_paper.tex
Update YSC paper

@@
 \DeclareMathOperator{\sign}{sign}
-\newtheorem{theorem}{Theorem}
+\newtheorem{hypothesis}{Hypothesis}
 
@@
 multidimensional problems to one or several one-dimensional ones, i. e. the \textit{dimension
-reduction} schemes have been developed \cite{stronginBook}, \cite{butzCurves}.
+reduction} schemes have been developed \cite{evolvents2013}, \cite{gergelGrishaginStringin2013}.
+All these schemes can be efficiently parallelized \cite{optParallelBook}.
 \par
@@
 nested optimization scheme has been considered. For each interval, the local estimate of the
-constant is an additive convolution of the ''global'' and ''local'' ones (\(\gamma\) and \(\lambda\), correspondingly):
+constant is an additive convolution of the ``global'' and ``local'' ones (\(\gamma\) and \(\lambda\), correspondingly):
 \begin{displaymath}
@@
 local minima, then the initial value of \(r\) is specified to be large enough that
-results in the dominance of the ''global'' component \(\gamma\) in the adaptive convolution.
+results in the dominance of the ``global'' component \(\gamma\) in the adaptive convolution.
 
 %------------------------------------------------------------------------------
-\section{Conditions of Convergence}
+\section{Hypothesis about Convergence}
 %\label{sect:algorithm}
@@
 function is Lipschitz one is given. However, as a rule, such statements are true in
-Hölder metrics as well, therefore, presumably, the following theorem is true:
+Hölder metrics as well, therefore, presumably, the following hypothesis is true:
 
-\begin{theorem}
+\begin{hypothesis}
 Assume the objective function \(f(x)\) to  satisfy Hölder condition with finite
@@
 \end{enumerate}
-\end{theorem}
+\end{hypothesis}
 
-The proof of this theorem requires further theoretical studies. Within the frames of
+The proof of this hypothesis requires further theoretical studies. Within the frames of
 the present work, it is not performed. The convergence has been established numerically only.
@@
 This research was supported by the Russian Science Foundation, project No 16-11-10150
-“Novel efficient methods and software tools for the time consuming decision making problems
-with using supercomputers of superior performance”.
+``Novel efficient methods and software tools for the time consuming decision making problems
+with using supercomputers of superior performance''.
 

### 2016-09-19 2016/YSC/ysc_paper.tex
Update conclusion in YSC paper

@@
 \section{Conclusions}
-In the present paper, the application of the methods of accounting for the local
-properties of the objective function
-in the multidimensional multiextremal optimization problems has been considered. Taking
-into account the local properties is expressed in the use of different estimates of Hölder
-constant in different search domains. The efficiency of the considered approach has been
-confirmed by solving the series of multidimensional problems form two classes.
+In the present work, the global optimization problems and the deterministic methods
+for solving these ones are considered. The relevance of solving the problems of this
+class is defined by the importance of the applications (optimal design, problems of
+parameter identification, pattern recognition, signa processing, etc.). The multiextremal
+(global) optimization methods are extremely computation costly since these ones require
+the constructing and using of the adaptive models of the objective function behavior
+according to the trial results accumulated during the algorithm execution.
+\par
+In this work, the application of the method of accounting for the local behavior of
+the optimized function within the multidimensional multiextremal optimization method
+has been considered. Earlier, this method has been applied to the one-dimensional
+problems only. Taking into account the local properties is implemented in the use of
+different estimates of Hölder constant in different search domains. Thus, the process
+of solving the problem is accelerated significantly by means of reducing the number of
+the search trials.
+\par
+The comparison of two different schemes of constructing the estimates of Hölder constant
+(\ref{additiveConv}), (\ref{additiveAdaptiveConv}) has revealed the advantage of scheme (\ref{additiveConv}).
+However, a larger value of the reliability parameter of the search method \(r\) might be required
+to use this scheme for solving some problems. When solving the complex problems with a small
+area of attraction of the global extremum, scheme (\ref{additiveAdaptiveConv}) is preferable for
+more reliability of the method. The efficiency of the considered approach has been confirmed
+by solving the series of multidimensional problems form two classes.
 

### 2016-09-22 2016/YSC/ysc_paper.tex
Update YSC paper

--- 2016-09-18 2016/YSC/ysc_paper.tex
@@
 \begin{abstract}
-  In the present paper, one of the methods of accounting for the local information on
-  the objective function in the Lipschitz global optimization problems has been considered.
-  In the course of solving such problems, a problem of estimating Lipschitz constant of the
-  objective function arises. According to the classic scheme, such an estimate is a single one
-  for the whole search domain. The method of accounting for the local properties is based on
-  the building of the estimates of Lipschitz constants for the search subdomains and has been
-  investigated earlier for the one-dimensional case. For solving the multidimensional problems,
-  the dimension reduction is applied. A multidimensional optimization problem is reduced to
-  a one-dimensional problem, which the objective function satisfy Hölder condition in.
-  In this paper, the application of building the local estimates of Hölder constant in the
-  reduced multidimensional optimization problems using the same scheme to the one used
-  in the one-dimensional Lipschitzian problems is considered.
+This paper considers one of the methods to account for the local information on the
+objective function in  Lipschitz global optimization problems. In the course of solving
+such problems, an issue arises regarding estimating the Lipschitz constant of the
+objective function arises. According to the classic scheme, this constant is a single
+estimate for the whole search domain. The method of accounting for local properties
+is based on building estimates of Lipschitz constants for the search subdomains,
+and has previously been investigated for the one-dimensional case. For solving
+multidimensional problems, dimension reduction is applied. A multidimensional
+optimization problem is reduced to a one-dimensional problem, for which the objective
+function satisfies the Hölder condition. In this paper, the application of building
+local Hölder constant estimates in reduced multidimensional optimization problems using
+the same scheme as is used in one-dimensional Lipschitzian problems is considered.
 \end{abstract}
@@
 
-The one-dimensional characteristical algorithms represent one of the classes of the
-optimization algorithms used widely \cite{optHandbook}. A number of schemes of reduction of the
-multidimensional problems to one or several one-dimensional ones, i. e. the \textit{dimension
-reduction} schemes have been developed \cite{evolvents2013}, \cite{gergelGrishaginStringin2013}.
+One-dimensional-characteristical algorithms represent one class of widely-used
+optimization algorithms \cite{optHandbook}. A number of schemes to reduce multidimensional problems
+to one or more one-dimensional problems, i.e. the \textit{dimension-reduction} schemes
+have been developed \cite{evolvents2013}, \cite{gergelGrishaginStringin2013}.
 All these schemes can be efficiently parallelized \cite{optParallelBook}.
 \par
-In recent years, when building the global optimization algorithms, an increased attention
-is paid to accounting for the local properties of the objective function. In the
-case of the characteristical schemes, this approach allows developing a mixed algorithm
-\cite{mixedAlg} as well as using different estimates of Lipschitz constant in different search
-domain \cite{sergLocalTuningFirst}. Originally, the latter approach has been proposed for the one-dimensional
-problems, also the investigation of this one when using the nested optimization scheme
-has been carried out \cite{nestedLocal}. In all these cases, this approach have demonstrated itself to
-be an efficient one, having allowed accelerating the convergence of the
-optimization methods essentially. In the present paper, a generalization of the method
-for building of the local estimates of Lipschitz constants onto the case of Hölder
-functions obtained when reducing the multidimensional optimization problems is considered.
+In recent years, when building global optimization algorithms, increased attention
+has been paid to accounting for the local properties of the objective function.
+In the case of characteristical schemes, this approach allows a mixed algorithm
+to be developed \cite{mixedAlg} and different estimates of Lipschitz constants to
+be used in different search domains \cite{sergLocalTuningFirst}. Originally, the
+latter approach had been proposed for one-dimensional problems, and this has also
+been examined when carrying out a nested optimization scheme \cite{nestedLocal}.
+In all these cases, this approach has proven to be efficient, having essentially
+accelerated the convergence of the optimization methods. This paper considers generalizing
+the method for building local estimates of Lipschitz constants in the case of Hölder
+functions obtained when reducing multidimensional optimization problems.
 
@@
 
-The global optimization problem can be formulated as follows:
-to find the global minimum of an \(N\)-dimensional function \(\varphi(y)\) in a hyperinterval
+The global optimization problem can be formulated as follows: to find the global minimum of an
+\(N\)-dimensional function \(\varphi(y)\) in a hyperinterval
 \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\): \(\varphi(y^*)=\min\{\varphi(y):y\in D\}\).
-In order to obtain an estimate of the global minimum from a finite number of computations
-of the function values, \(\varphi(y)\) is required to satisfy Lipschitz condition.
+In order to obtain an estimate of the global minimum from a finite number of function value computations,
+ \(\varphi(y)\) is required to satisfy Lipschitz condition.
 \begin{displaymath}
@@
 \par
-The use of the evolvents \(y(x)\) i. e. the curves filling the space is a classic scheme
-for the dimension reduction for the global optimization algorithms \cite{strOptBook}.
+The use of the evolvents \(y(x)\) i.e. the curves filling the space are a classic
+dimension-reduction scheme for global optimization algorithms \cite{strOptBook}.
 \begin{displaymath}
@@
 \par
-Such a mapping allows reducing a problem stated in a multidimensional space to solving
-a one-dimensional one at the expense of worsening its properties. In particular,
-the one-dimensional function \(\varphi(y(x))\) is not a Lipschitz one but is a Hölder one:
+Such a mapping allows the reduction of a problem stated in a multidimensional space
+to solving a one-dimensional problem at the expense of degrading its properties.
+In particular, the one-dimensional function \(\varphi(y(x))\) is not a Lipschitzian but a Hölderian function:
 \begin{displaymath}
@@
 \end{displaymath}
-where Hölder constant \(H\) is related to Lipschitz constant \(L\) by the relation
+where the Hölder constant \(H\) is related to the Lipschitz constant \(L\) by the relation
 \begin{displaymath}
@@
 Therefore, not limiting the generality, one can consider the minimization of the
-one-dimensional function \(f(x)=\varphi(y(x))\), \(x\in[0;1]\) satisfying Hölder condition.
-The issues of numerical building of the mapping like Peano curve and the corresponding
-theory have been considered in details in \cite{strOptBook}. Here we would note that an evolvent
-built numerically is an approximation to the theoretical Peano curve with the precision
+one-dimensional function \(f(x)=\varphi(y(x))\), with \(x\in[0;1]\) satisfying Hölder condition.
+The issues of numerically building the mapping like a Peano curve and the corresponding
+theory have been considered in detail in \cite{strOptBook}. Here we would note that an evolvent
+built numerically is an approximation to the theoretical Peano curve with a precision
 of the order \(2^{-m}\) where \(m\) is the building parameter of the evolvent.
@@
 
-The considered algorithm of solving the stated problem implies the building of
-a sequence of points \(x_k\), which the values of the minimized function \(z_k = f(x_k)\)
-are computed in. Let us call the process of computation of the function value
-(including the building of an image \(y_k=y(x_k)\)) a trial, and the pair \((x_k,z_k)\) ---
+The algorithm considered for solving the stated problem implies building
+a sequence of points \(x_k\), in which the values of the minimized function \(z_k = f(x_k)\)
+are computed. Let us call the process of computating the function value
+(including building an image \(y_k=y(x_k)\)) a trial, and the pair \((x_k,z_k)\) ---
 the result of the trial. A set of the pairs \(\{(x_k,z_k)\}, 1\leqslant k\leqslant n\)
-makes the search information accumulated by the method after executing \(n\) steps.
+makes up the search information accumulated by the method after executing \(n\) steps.
 \par
-At the first iteration of the method, a trial is executed in an arbitrary internal
-point \(x_1\) of the interval \([0,1]\). Let us assume \(n \geqslant 1\) iterations
-of the method to be completed, in the course of which the iterations in \(k = k(n)\)
+At the first iteration of the method, a trial is executed at an arbitrary internal
+point \(x_1\) within the interval \([0,1]\). Let us assume \(n \geqslant 1\) iterations
+of the method to be completed, during the course of which the iterations in \(k = k(n)\)
 points \(x_i, 1\leqslant i\leqslant k\) have been performed. Then, the point \(x^{k+1}\)
-of the search trial of the next \((k+1)\)th iteration is determined in accordance with the rules:
+of the search trial for the next \((k+1)\)th iteration is determined in accordance with the rules:
 \par
-Step 1. Renumber the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\),
+Step 1. Renumber the points in the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\),
 which includes the boundary points of the interval \([0,1]\) as well as the points of
-preceding trials, by the lower indices in the order of increasing the values of coordinate i. e.
+preceding trials, by the lower indices in order of increasing coordinate values  i.e.
 \begin{displaymath}
@@
 \end{equation}
-where \(r\) is a predefined parameter of the method, and \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
+where \(r>1\) is a predefined parameter for the method, and \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
 \par
@@
 \par
-Step 5. Execute new trial in the point \(x_{k+1}\) computed according to the formulae
+Step 5. Execute a new trial at point \(x_{k+1}\) computed according to the formulae
 \begin{displaymath}
@@
 The algorithm is terminated if the condition \(\Delta_{t}\leqslant \varepsilon\) is fulfilled;
-here \(\varepsilon>0\) is a predefined precision. As an estimate of the global optimum solution of the problem the values
+here \(\varepsilon>0\) is the predefined precision. As an estimate of the global optimum solution of the problem the values
 \begin{equation}
@@
 %\label{sect:algorithm}
-At it is seen from the scheme of the algorithm, regardless to the local properties
-of the optimized one-dimensional function, the same values of estimate of Hölder
+As it is seen from the scheme of the algorithm, regardless of the local properties
+of the optimized one-dimensional function, the same estimate of the Hölder
 constant (\ref{step2}) is used to compute the characteristic of all intervals (\ref{step3_1}), (\ref{step3_2}).
 In \cite{sergLocalTuning}, it has been proposed to use different values of \(M\) for each interval,
-and also the efficiency of this approach in the case of the optimization of the
-one-dimensional functions satisfying Lipschitz condition has been shown. In \cite{nestedLocal},
+and also the efficiency of this approach has been shown in the case of optimizing
+one-dimensional functions that satisfy the Lipschitz condition. In \cite{nestedLocal},
 the application of the adaptive estimates of Lipschitz constants in the multidimensional
 nested optimization scheme has been considered. For each interval, the local estimate of the
-constant is an additive convolution of the ``global'' and ``local'' ones (\(\gamma\) and \(\lambda\), correspondingly):
+constant is an additive convolution of the ``global'' and ``local'' constants
+(\(\gamma\) and \(\lambda\), correspondingly):
 \begin{displaymath}
@@
 \end{equation}
-\(\xi\) is chosen small to prevent the case of the function being constant over search interval.
+A small \(\xi\) is chosen to preventthe function from being constant over the search interval.
 This variant of convolution does not depend on the parameter \(r\), however, the adaptive convolution:
@@
 If it is known a priori that the optimized function has a complex shape with multiple
-local minima, then the initial value of \(r\) is specified to be large enough that
-results in the dominance of the ``global'' component \(\gamma\) in the adaptive convolution.
+local minima, then the initial value of \(r\) is specified to be large enough to
+result in the dominance of the ``global'' component \(\gamma\) in the adaptive convolution.
 
@@
 %\label{sect:algorithm}
-In \cite{sergLocalTuning}, a theorem on the convergence of the method in the case if the objective
-function is Lipschitz one is given. However, as a rule, such statements are true in
-Hölder metrics as well, therefore, presumably, the following hypothesis is true:
+In \cite{sergLocalTuning}, a theorem is given on the convergence of the method in the
+case of a Lipschitz objective function. However, as a rule, such statements are
+true for Hölder metrics as well. Therefore, the following hypothesis is presumably true:
 
@@
 \begin{enumerate}
-  \item If \(x\in(0;1)\), then convergence to \(x\) is a bilateral one i. e. there
-  exist two infinite subsequences of \(\{x_k\}\) converging to \(x\): one from the
+  \item If \(x\in(0;1)\), then convergence to \(x\) is a bilateral one i.e. there
+  exists two infinite subsequences of \(\{x_k\}\) converging to \(x\): one from the
   left, the other from the right;
@@
   \item If the function \(f(x)\) has a finite number of local minima in \([0, 1]\),
-  then the point \(x\) is a local optimum one;
+  then the point \(x\) is a local optimum;
   \item (Sufficient conditions for convergence to a global minimizer). Let \(x^*\)
@@
   such that for all \(k > k^*\) the inequality
-  \(M_j(k) > H_j(k)\) holds, where \(H_j(k)\) is Hölder constant for the interval
+  \(M_j(k) > H_j(k)\) holds, where \(H_j(k)\) is the Hölder constant for the interval
   \([x_{j(k)-1}, x_{j(k)}]\) containing \(x^*\), and \(M_{j(k)}\) is its estimate.
-  Then, the set of limit points of the sequence \(\{x_k\}\) coincides with the set
-  of global minimizers of the function \(f(x)\).
+  Then, the set of limit points for the sequence \(\{x_k\}\) coincides with the set
+  of global minimizers for the function \(f(x)\).
 \end{enumerate}
@@
 
-The proof of this hypothesis requires further theoretical studies. Within the frames of
-the present work, it is not performed. The convergence has been established numerically only.
+The proof of this hypothesis requires further theoretical studies. It was not performed
+as part of this present work. The convergence has only been established numerically.
 
 %------------------------------------------------------------------------------
-\section{Results of Experiments}
+\section{Experimental Results}
 \label{sect:experiments}
 
-The experiments on the evaluation of the efficiency of the method with the local
-adaptive estimate of Hölder constant have been carried out using the two-dimensional
-classes of Grishagin problems (\(F_{GR}\)) \cite{grishaginClass} and GKLS ones \cite{gklsClass}. Each class
-includes 100 multiextremal functions. In all experiments, the evolvent was built
-with the density \(m = 12\), parameter \(\varepsilon\) in the termination criterion was equal \(10^{-3}\).
-The parameter \(r\) was selected as low as possible, allowing to solve all the
-problems. The search step for \(r\) was \(0.1\).
+The experiments to evaluate the efficiency of the method with a local adaptive
+estimate of the Hölder constant have been carried out using the two-dimensional
+classes of Grishagin (\(F_{GR}\)) \cite{grishaginClass} and GKLS problems \cite{gklsClass}.
+Each class includes 100 multiextremal functions. In all experiments, the evolvent was
+built with the density \(m = 12\), the parameter \(\varepsilon\) in the termination criterion
+was set to \(10^{-3}\). The parameter \(r\) was selected as low as possible, at which
+the given method solves all problems of the class. The search step for \(r\) was \(0.1\).
+
 \par
-For clarity of illustration of the advantages of the local adaptive scheme of estimating
+For clarity in illustrating the advantages of the local adaptive scheme for estimating
 the constant \(H\), let us consider a particular example of the results obtained by the method.
-In Fig. \ref{fig:grish_isolines}, the level lines of a function from \(F_{GR}\) class and the trial
-points executed by the method with the global estimate of Hölder constant and with
-the one estimated according to formula (\ref{additiveConv}) are shown. As it is
-seen from the figures, the method with global estimate of the constant executes a
-large number of trials in the nearness of the global minimizer (total 1086 trials
-have been executed) before the termination condition is satisfied, whereas the method
-with the local adaptive estimate converges much faster (total 385 trials have been executed).
-The same situation takes place when optimizing a function from GKLS class (Fig. \ref{fig:gkls_isolines}).
-The method with the global estimate of the constant had executed 2600 trials while
-the method with the local adaptive estimate had executed 1190 trials only.
+In Figure \ref{fig:grish_isolines}, the level lines of a function from the \(F_{GR}\) class and
+ the trial points executed by the method with a global estimate of the Hölder constant, and with
+ the one estimated according to formula (\ref{additiveConv}) are shown. As can be seen from
+ the figures, the method using a global estimate of the constant executed a large
+ number of trials in the vicinity of the global minimizer (a total of 1,086 trials
+ were executed) before the termination condition was satisfied, whereas the method
+ with the local adaptive estimate converged much faster (a total of 385 trials
+ were executed). The same situation occurs when optimizing a function from the
+ GKLS class (Fig. \ref{fig:gkls_isolines}). The method using the global estimate of
+ the constant executed 2,600 trials while the method using the local adaptive estimate only
+ executed 1,190 trials.
+
 \begin{figure}[ht]
@@
     \subfloat[local-adaptive \(H\) estimation]{{\includegraphics[width=0.45\textwidth]{images/gkls_loc.png} }}
-    \caption{The contour plots of a function from GKLS class}
+    \caption{The level lines of a function from GKLS class}
     \label{fig:gkls_isolines}
@@
 \par
-Next, let us go to the comparison of various variants of the method for the problem
-classes considered. In order to evaluate the efficiency of an algorithm, we will use
-the operating characteristics \cite{grishaginClass}, which is defined by a set of points on the \((K, P)\)
-plane where \(K\) is the averaged number of the search trials carried out before the satisfaction
-of the termination condition when minimizing a function from given class, and \(P\)
-is portion of problems solved successfully. If at given \(K\) the operating characteristic
-of a method goes higher than the one of another method, it means that at fixed costs
-of the search the former method will find the solution with a greater probability.
-If some value of \(P\) is fixed, and the characteristic of a method goes to the left from
-the one of another method, the former method requires fewer resources to achieve the same reliability.
-\par
+Next, let’s compare the various alternatives for the method in the problem classes
+considered. In order to evaluate the efficiency of an algorithm, we will use the
+operating characteristics \cite{grishaginClass}, which are defined by a set of
+points on the \((K, P)\) plane where \(K\) is the average number of search trials
+conducted before satisfying the termination condition when minimizing a function
+from a given class, and \(P\) is the proportion of problems solved successfully.
+If at a given \(K\), the operating characteristic of a method goes higher than one
+from another method, it means that at fixed search costs, the former method has a
+greater probability of finding the solution. If some value of \(P\) is fixed, and the
+characteristic of a method goes to the left from that of another method, the former
+method requires fewer resources to achieve the same reliability.
+
 \begin{figure}[ht]
@@
 \end{figure}
-As it is seen from the operating characteristics in Fig. \ref{fig:grishh_op}, \ref{fig:gkls_op}, both methods with the
-local adaptive estimate of Hölder constant have demonstrated an essential advantage.
-However, both methods require setting a higher value of the parameter \(r\). If one compares the
-methods with the adaptive convolution and  with the non-adaptive one (\ref{additiveConv})(\ref{additiveAdaptiveConv}), the advantage
-of the latter is seen clearly, although it requires a higher value of the reliability
-parameter \(r\) to solve the problems form GKLS class.
+\par
+As can be seen from the operating characteristics in Figures \ref{fig:grishh_op}
+and \ref{fig:gkls_op}, both methods using a local adaptive estimate of the Hölder
+constant have demonstrated an essential advantage. However, both methods require setting
+a higher value for the parameter \(r\). If one compares the methods with both the
+adaptive and non-adaptive convolution (\ref{additiveConv})(\ref{additiveAdaptiveConv}),
+the advantage of the latter can clearly be seen, although it requires a higher value
+for the reliability parameter \(r\) to solve problems from the GKLS class.
+
   \begin{figure}[ht]
@@
 \section{Conclusions}
-In the present paper, the application of the methods of accounting for the local
-properties of the objective function
-in the multidimensional multiextremal optimization problems has been considered. Taking
-into account the local properties is expressed in the use of different estimates of Hölder
-constant in different search domains. The efficiency of the considered approach has been
-confirmed by solving the series of multidimensional problems form two classes.
+This present work considers global optimization problems and the deterministic methods
+for solving them. The relevance of solving this class of problems is defined by the
+importance of the applications (optimal design, problems of parameter identification,
+pattern recognition, signal processing, etc.). The multiextremal (global) optimization
+methods are extremely computationally costly, since these require constructing and using
+adaptive models of the objective function behavior according to the trial results
+accumulated while executing the algorithm.
+\par
+This paper considers the application of a method to account for the local behavior
+of the objective function within the multidimensional multiextremal optimization method.
+This method has previously been applied to one-dimensional problems only. Taking into
+account the local properties are implemented in using different estimates of Hölder
+constant within different search domains. Thus, the process of solving the problem is
+accelerated significantly by means of reducing the number of search trials.
+\par
+The comparison of two different schemes for constructing estimates of the Hölder
+constant (\ref{additiveConv}), (\ref{additiveAdaptiveConv}) has revealed the advantage
+of scheme (\ref{additiveConv}). However, a larger value for the reliability parameter \(r\)
+in the search method might be required to use this scheme for solving certain problems.
+When solving complex problems with a small area of attraction for the global extremum,
+scheme (\ref{additiveAdaptiveConv}) is preferable for the greater reliability of this method.
+The efficiency of the considered approach has been confirmed by solving the series of
+multidimensional problems from two classes.
 
@@
 This research was supported by the Russian Science Foundation, project No 16-11-10150
-``Novel efficient methods and software tools for the time consuming decision making problems
-with using supercomputers of superior performance''.
+``Novel efficient methods and software tools for the time consuming decision- making problems
+with using supercomputers of superior performance.''
 

### 2016-09-25 2016/YSC/ysc_paper.tex
Update YSC paper

@@
 \(N\)-dimensional function \(\varphi(y)\) in a hyperinterval
-\(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\): \(\varphi(y^*)=\min\{\varphi(y):y\in D\}\).
+\(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\):
+\begin{displaymath}
+  \varphi(y^*)=\min\{\varphi(y):y\in D\}.
+\end{displaymath}
 In order to obtain an estimate of the global minimum from a finite number of function value computations,
@@
 The use of the evolvents \(y(x)\) i.e. the curves filling the space are a classic
-dimension-reduction scheme for global optimization algorithms \cite{strOptBook}.
+dimension-reduction scheme for global optimization algorithms \cite{evolvents2013}.
 \begin{displaymath}
@@
 Such a mapping allows the reduction of a problem stated in a multidimensional space
-to solving a one-dimensional problem at the expense of degrading its properties.
+to solving a one-dimensional problem at the expense of worsening its properties.
 In particular, the one-dimensional function \(\varphi(y(x))\) is not a Lipschitzian but a Hölderian function:
@@
 Therefore, not limiting the generality, one can consider the minimization of the
-one-dimensional function \(f(x)=\varphi(y(x))\), with \(x\in[0;1]\) satisfying Hölder condition.
+one-dimensional function \(f(x)=\varphi(y(x))\), with \(x\in[0,1]\) satisfying Hölder condition.
 The issues of numerically building the mapping like a Peano curve and the corresponding
-theory have been considered in detail in \cite{strOptBook}. Here we would note that an evolvent
+theory have been considered in detail in \cite{evolvents2013}. Here we would note that an evolvent
 built numerically is an approximation to the theoretical Peano curve with a precision
@@
 
-The algorithm considered for solving the stated problem implies building
+The algorithm considered for solving the stated problem implies generating
 a sequence of points \(x_k\), in which the values of the minimized function \(z_k = f(x_k)\)
 are computed. Let us call the process of computating the function value
-(including building an image \(y_k=y(x_k)\)) a trial, and the pair \((x_k,z_k)\) ---
-the result of the trial. A set of the pairs \(\{(x_k,z_k)\}, 1\leqslant k\leqslant n\)
+(including calulating an image \(y^k=y(x^k)\)) a trial, and the pair \((x^k,z^k)\) ---
+the result of the trial. A set of the pairs \(\{(x^k,z^k)\}, 1\leqslant k\leqslant n\)
 makes up the search information accumulated by the method after executing \(n\) steps.
@@
 At the first iteration of the method, a trial is executed at an arbitrary internal
-point \(x_1\) within the interval \([0,1]\). Let us assume \(n \geqslant 1\) iterations
+point \(x^1\) within the interval \([0,1]\). Let us assume \(n \geqslant 1\) iterations
 of the method to be completed, during the course of which the iterations in \(k = k(n)\)
@@
 \par
-Step 2. Assuming \(z_i=f(x_i),1\leqslant i\leqslant k\), compute the quantities
+Step 2. Assuming \(z_i=f(x_i),1\leqslant i\leqslant k\), compute the values
 \begin{equation}
@@
 \label{step3_1}
-R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M}
+R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
 \end{equation}
@@
 \label{step3_2}
-R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1
+R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1.
 \end{equation}
@@
 \begin{displaymath}
-x_{k+1}=\dfrac{x_{t}+x_{t-1}}{2},t=1,t=k+1
+x_{k+1}=\dfrac{x_{t}+x_{t-1}}{2},t=1,t=k+1,
 \end{displaymath}
@@
 \label{step5}
-x_{k+1}=\dfrac{x_{t}+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{\mu}\right]^N,1<t<k+1
+x_{k+1}=\dfrac{x_{t}+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{\mu}\right]^N,1<t<k+1.
 \end{equation}
@@
 The algorithm is terminated if the condition \(\Delta_{t}\leqslant \varepsilon\) is fulfilled;
-here \(\varepsilon>0\) is the predefined precision. As an estimate of the global optimum solution of the problem the values
+here \(\varepsilon>0\) is the predefined accuracy. As an estimate of the global optimum solution of the problem the values
 \begin{equation}
@@
 \end{equation}
-are selected. The theoretical substantiation of this method is presented in \cite{strOptBook}.
+are selected. The theoretical substantiation of this method is presented in \cite{strOptBook}, chapter 8.
 
@@
 constant (\ref{step2}) is used to compute the characteristic of all intervals (\ref{step3_1}), (\ref{step3_2}).
-In \cite{sergLocalTuning}, it has been proposed to use different values of \(M\) for each interval,
+In \cite{sergLocalTuningFirst}, it has been proposed to use different values of \(M\) (\(M\) is the same as in (\ref{step2}), (\ref{step3_1})) for each interval,
 and also the efficiency of this approach has been shown in the case of optimizing
@@
 \end{equation}
-A small \(\xi\) is chosen to preventthe function from being constant over the search interval.
+A small \(\xi\) is chosen to prevent the function from being constant over the search interval.
 This variant of convolution does not depend on the parameter \(r\), however, the adaptive convolution:
@@
 \end{equation}
-has been considered in \cite{sergLocalTuning} as well.
+has been introduced in \cite{gergel2010} and detailed considered (in case of solving one-dimentional Lipschitzian problems) in \cite{sergLocalTuning} as well.
 If it is known a priori that the optimized function has a complex shape with multiple

### 2016-11-27 2017/PAVT/examin-poster.tex
Add tex variant of the PAVT2017 poster introduction

\documentclass[11pt, oneside, a4paper]{article}
\usepackage[utf8]{inputenc}
%\usepackage[cp1251]{inputenc} % кодировка
\usepackage[english, russian]{babel} % Русские и английские переносы
\usepackage{graphicx}          % для включения графических изображений
\usepackage{cite}              % для корректного оформления литературы
\usepackage{pavt-ru}

\begin{document}

% \title - название статьи
% \authors - список авторов

\title{Использование параллельной системы глобальной оптимизации Globalizer для
решения задач оптимального управления\footnote{Исследование выполнено при финансовой
поддержке РФФИ в рамках научного проекта № 16-31-00244 мол\_а <<Параллельные методы
решения вычислительно трудоемких задач глобальной оптимизации на гибридных кластерных системах>>}}

\authors{И.Г.~Лебедев, В.В.~Соврасов}
\organizations{Нижегородский государственный университет им. Н.И. Лобачевского}

Задача многомерной многоэкстремальной оптимизации может быть определена как проблема
поиска наименьшего значения действительной функции \(\varphi(y)\) в некоторой области \(D\),
задаваемой функциональными ограничениями:
\begin{displaymath}
  \label{task}
    \varphi(y^*)=\min\{\varphi(y):y\in D\}
\end{displaymath}
\begin{displaymath}
  D=\{x\in \mathbf{R}^n: g_j(x) \le 0, j=\overline{1,m}\}
\end{displaymath}

В ННГУ им. Н.И. Лобачевского под руководством проф. Р.Г. Стронгина разработан
эффективный подход к решению задач глобальной оптимизации \cite{strGergrParOptBook}.
В рамках данного подхода решение многомерных задач сводится к решению серии вложенных задач меньшей размерности.
Главным из рассматриваемых способов редукции размерности является использование кривой Пеано,
однозначно отображающей отрезок вещественной оси \([0,1]\) на \(n\)-мерный куб.
Для организации параллельных вычислений используется параллельный алгоритм глобального поиска,
эффективность которого была показана ранее в \cite{parallelMethod}.

Задача глобальной оптимизации возникает при синтезе оптимальных с точки зрения некоторых критериев управлений в линейных системах ОДУ.
Если управление является линейной обратной связи по состоянию, то система с управлением имеет вид:
\begin{displaymath}
    \dot x = (A+B_u\Theta)x + B_v v, x(0)=0,
\end{displaymath}
где  \(v(t)\in L_2\) --- некоторое возмущение.
Выходы системы описываются формулами \(z_k=(C_k+B_u\Theta),k=\overline{1,N}\).
Вляиние возмущения на \(k\)-й выход системы описывается критерием:
\begin{displaymath}
  J_k(\Theta)=\sup_{v\in L_2} \frac{\max_{1\le i \le n_k} \sup_{t\ge 0}|z_k^{(i)}(t)|}{||v||_2}
\end{displaymath}

Задача оптимизации состоит в том, чтобы найти компоненты вектора \(\Theta\), минимизирующие
один из критериев при заданных ограничениях на другие: \(J_1(\Theta^*)=\min\{J_1(\Theta):J_k(\Theta)\le S_k,k=\overline{2,N}\})\).

В \cite{optControl} указан способ вычисления критериев, сводящийся к решении нескольких задач линейной алгебры,
а также даны решения задачи поиска оптимального управления в некоторых частных случаях. На данных момент результаты,
полученные в \cite{optControl} повторены с помощью системы Globalizer, ведётся
подготовка к решению более сложных задач из рассматриваемого класса, в которых вычисление критериев --- довольно
трудоемкий процесс, требующий привлечения ресурсов вычислительного кластера.

\begin{biblio}

\bibitem{strGergrParOptBook}
Стронгин Р.Г. Гергель В.П. Гришагин
  В.А.~Баркалов К.А.
\newblock {\em Параллельные вычисления в задачах
  глобальной оптимизации}.
\newblock М.: Издательство Московского
  университета, 2013, 280с.

 \bibitem{parallelMethod}
 И.Г. Лебедев, В.В. Соврасов

\bibitem{optControl}
 Д.В. Баландин М.М. Коган
 \newblock {\em Оптимальное по Парето обобщенное
 \(H_2\)-управление и задачи виброзащиты}.

\end{biblio}
\end{document}

### 2016-11-28 2017/PAVT/examin-poster.tex
Update PAVT poster

@@
   \label{task}
-    \varphi(y^*)=\min\{\varphi(y):y\in D\}
-\end{displaymath}
-\begin{displaymath}
-  D=\{x\in \mathbf{R}^n: g_j(x) \le 0, j=\overline{1,m}\}
+    \varphi(y^*)=\min\{\varphi(y):y\in D\}, D=\{x\in \mathbf{R}^n: g_j(x) \le 0, j=\overline{1,m}\}
 \end{displaymath}
@@
 эффективный подход к решению задач глобальной оптимизации \cite{strGergrParOptBook}.
-В рамках данного подхода решение многомерных задач сводится к решению серии вложенных задач меньшей размерности.
-Главным из рассматриваемых способов редукции размерности является использование кривой Пеано,
-однозначно отображающей отрезок вещественной оси \([0,1]\) на \(n\)-мерный куб.
-Для организации параллельных вычислений используется параллельный алгоритм глобального поиска,
-эффективность которого была показана ранее в \cite{parallelMethod}.
+В рамках данного подхода решение многомерной задачи сводится к решению одномерной.
+Для редукции размерности используются кривые Пеано, однозначно отображающие отрезок
+вещественной оси \([0,1]\) на \(n\)-мерный куб. Для организации параллельных вычислений
+используется параллельный алгоритм глобального поиска, эффективность которого была
+показана ранее в \cite{parallelMethod}.
 
-Задача глобальной оптимизации возникает при синтезе оптимальных с точки зрения некоторых критериев управлений в линейных системах ОДУ.
-Если управление является линейной обратной связи по состоянию, то система с управлением имеет вид:
+Задача глобальной оптимизации возникает при синтезе оптимальных с точки зрения некоторых
+критериев управлений в линейных системах ОДУ. Если управление является линейной обратной
+связью по состоянию, то система с управлением имеет вид:
 \begin{displaymath}
@@
 \begin{displaymath}
-  J_k(\Theta)=\sup_{v\in L_2} \frac{\max_{1\le i \le n_k} \sup_{t\ge 0}|z_k^{(i)}(t)|}{||v||_2}
+  J_k(\Theta)=\sup_{v\in L_2} \frac{\max_{1\le i \le n_k} \sup_{t\ge 0}|z_k^{(i)}(\Theta,t)|}{||v||_2}
 \end{displaymath}
 
-Задача оптимизации состоит в том, чтобы найти компоненты вектора \(\Theta\), минимизирующие
-один из критериев при заданных ограничениях на другие: \(J_1(\Theta^*)=\min\{J_1(\Theta):J_k(\Theta)\le S_k,k=\overline{2,N}\})\).
+Нужно найти компоненты вектора \(\Theta\), минимизирующие один из критериев при
+заданных ограничениях на другие: \(J_1(\Theta^*)=\min\{J_1(\Theta):J_k(\Theta)\le S_k,k=\overline{2,N}\}\).
 
-В \cite{optControl} указан способ вычисления критериев, сводящийся к решении нескольких задач линейной алгебры,
-а также даны решения задачи поиска оптимального управления в некоторых частных случаях. На данных момент результаты,
-полученные в \cite{optControl} повторены с помощью системы Globalizer, ведётся
-подготовка к решению более сложных задач из рассматриваемого класса, в которых вычисление критериев --- довольно
-трудоемкий процесс, требующий привлечения ресурсов вычислительного кластера.
+В \cite{optControl} указан способ вычисления критериев, состоящий в решении задач линейной алгебры,
+а также даны решения рассматриваемой задачи в некоторых частных случаях. На
+данных момент результаты, полученные в \cite{optControl} повторены с помощью системы Globalizer,
+ведётся подготовка к решению более сложных задач из рассматриваемого класса, в которых
+вычисление критериев --- трудоемкий процесс, требующий привлечения ресурсов вычислительного кластера.
 
@@
   В.А.~Баркалов К.А.
-\newblock {\em Параллельные вычисления в задачах
+\newblock {Параллельные вычисления в задачах
   глобальной оптимизации}.
@@
  \bibitem{parallelMethod}
- И.Г. Лебедев, В.В. Соврасов
+Баркалов К.А. Лебедев И.Г. Соврасов В.В. Сысоев А.В.
+\newblock {Реализация параллельного
+алгоритма поиска глобального экстремума функции на Intel XEON PHI //
+Параллельные вычислительные технологии (ПаВТ’2016) труды международной
+научной конференции. 2016. С. 68-80.}
 
@@
  Д.В. Баландин М.М. Коган
- \newblock {\em Оптимальное по Парето обобщенное
+ \newblock {Оптимальное по Парето обобщенное
  \(H_2\)-управление и задачи виброзащиты}.

### 2016-11-28 2017/PAVT/examin-poster.tex
Fit PAVT poster to one page

@@
 Выходы системы описываются формулами \(z_k=(C_k+B_u\Theta),k=\overline{1,N}\).
-Вляиние возмущения на \(k\)-й выход системы описывается критерием:
-\begin{displaymath}
-  J_k(\Theta)=\sup_{v\in L_2} \frac{\max_{1\le i \le n_k} \sup_{t\ge 0}|z_k^{(i)}(\Theta,t)|}{||v||_2}
-\end{displaymath}
+Вляиние возмущения на \(k\)-й выход системы описывается критерием
+\(  J_k(\Theta)=\sup_{v\in L_2} \frac{\max_{1\le i \le n_k} \sup_{t\ge 0}|z_k^{(i)}(\Theta,t)|}{||v||_2}\).
 
@@
 
-В \cite{optControl} указан способ вычисления критериев, состоящий в решении задач линейной алгебры,
+В \cite{optControl} указан способ вычисления критериев, состоящий в решении СЛАУ,
 а также даны решения рассматриваемой задачи в некоторых частных случаях. На
@@
  \newblock {Оптимальное по Парето обобщенное
- \(H_2\)-управление и задачи виброзащиты}.
+ \(H_2\)-управление и задачи виброзащиты. Автоматика и телемеханика, в печати.}.
 

### 2017-02-05 2017/PAVT/examin-poster.tex
Добавлены правки в описание плаката

@@
 а также даны решения рассматриваемой задачи в некоторых частных случаях. На
-данных момент результаты, полученные в \cite{optControl} повторены с помощью системы Globalizer,
-ведётся подготовка к решению более сложных задач из рассматриваемого класса, в которых
+данных момент результаты, полученные в \cite{optControl}, повторены с помощью системы Globalizer,
+ведётся подготовка к решению задач большей размерности из рассматриваемого класса, в которых
 вычисление критериев --- трудоемкий процесс, требующий привлечения ресурсов вычислительного кластера.
@@
  \newblock {Оптимальное по Парето обобщенное
- \(H_2\)-управление и задачи виброзащиты. Автоматика и телемеханика, в печати.}.
+ \(H_2\)-управление и задачи виброзащиты // Автоматика и телемеханика. Принято к печати.}
 

### 2017-02-05 2017/PAVT/examin-poster.tex
PAVT: update poster description

@@
 \organizations{Нижегородский государственный университет им. Н.И. Лобачевского}
+\setlength{\abovedisplayskip}{3pt}
+\setlength{\belowdisplayskip}{3pt}
 
-Задача многомерной многоэкстремальной оптимизации может быть определена как проблема
-поиска наименьшего значения действительной функции \(\varphi(y)\) в некоторой области \(D\),
+Задача многомерной многоэкстремальной оптимизации может быть поставлена следующим образом:
+необходимо найти наименьшее значение вещественной функкции \(\varphi(y)\) в некоторой области \(D\),
 задаваемой функциональными ограничениями:
@@
 Вляиние возмущения на \(k\)-й выход системы описывается критерием
-\(  J_k(\Theta)=\sup_{v\in L_2} \frac{\max_{1\le i \le n_k} \sup_{t\ge 0}|z_k^{(i)}(\Theta,t)|}{||v||_2}\).
+\(J_k(\Theta)=\sup_{v\in L_2} \frac{\max_{1\le i \le n_k} \sup_{t\ge 0}|z_k^{(i)}(\Theta,t)|}{||v||_2}\).
 

### 2017-02-09 2017/PAVT/examin-poster.tex
PAVT: improve formatting

@@
 \usepackage{pavt-ru}
+\usepackage{amssymb}
 
@@
   \label{task}
-    \varphi(y^*)=\min\{\varphi(y):y\in D\}, D=\{x\in \mathbf{R}^n: g_j(x) \le 0, j=\overline{1,m}\}
+    \varphi(y^*)=\min\{\varphi(y):y\in D\}, D=\{x\in \mathbf{R}^n: g_j(x) \leqslant 0, j=\overline{1,m}\}
 \end{displaymath}
@@
 Вляиние возмущения на \(k\)-й выход системы описывается критерием
-\(J_k(\Theta)=\sup_{v\in L_2} \frac{\max_{1\le i \le n_k} \sup_{t\ge 0}|z_k^{(i)}(\Theta,t)|}{||v||_2}\).
+\(J_k(\Theta)=\sup_{v\in L_2} \frac{\max_{1\leqslant i \leqslant n_k} \sup_{t\geqslant 0}|z_k^{(i)}(\Theta,t)|}{||v||_2}\).
 
 Нужно найти компоненты вектора \(\Theta\), минимизирующие один из критериев при
-заданных ограничениях на другие: \(J_1(\Theta^*)=\min\{J_1(\Theta):J_k(\Theta)\le S_k,k=\overline{2,N}\}\).
+заданных ограничениях на другие: \(J_1(\Theta^*)=\min\{J_1(\Theta):J_k(\Theta)\leqslant S_k,k=\overline{2,N}\}\).
 

### 2017-03-05 2017/PAVT/poster/poster.tex
PAVT: add a placeholder for poster

\documentclass{beamer}
\usepackage[orientation=portrait, size=a1, scale=1.4, debug]{beamerposter}
\usetheme{unn}

\usepackage{cmbright}

\usepackage{fontspec}
\setmainfont{CMU Sans Serif}
\setromanfont{CMU Sans Serif}
\setsansfont{CMU Sans Serif}

\usepackage{polyglossia}
\setmainlanguage{russian}

\usepackage{booktabs}
\usepackage{ragged2e}

\usepackage[backend=biber,
            movenames=false,
            maxnames=4,
            style=gost-numeric,
            sorting=nty,
            autolang=other]{biblatex}
\addbibresource{bibliography.bib}

\title{Использование параллельной системы глобальной оптимизации Globalizer для
решения задач оптимального управления}
\author{И.Г. Лебедев \and В. В. Соврасов}
\institute{ННГУ им. Н.И. Лобачевского}

\begin{document}
\begin{frame}[t]
    \begin{columns}[t]
        \begin{column}[t]{0.48\paperwidth}
            \begin{block}{Метод глобальной оптимизмции}
            \end{block}
            \begin{block}{Задача поиска оптимального управления}
            \end{block}
        \end{column}
        \begin{column}[t]{0.48\paperwidth}
            \begin{block}{Результаты}
            \end{block}
        \end{column}
    \end{columns}
\end{frame}
\end{document}

### 2017-03-05 2017/PAVT/poster/poster.tex
PAVT: add task description in poster

@@
 
+\DeclareMathOperator{\re}{\operatorname{Re}}
+
 \title{Использование параллельной системы глобальной оптимизации Globalizer для
@@
 \institute{ННГУ им. Н.И. Лобачевского}
+\setlength{\abovedisplayskip}{3pt}
+\setlength{\belowdisplayskip}{3pt}
 
@@
         \begin{column}[t]{0.48\paperwidth}
+            \begin{block}{Задача поиска оптимального управления}
+              Необходимо найти оптимальное управление в виде обратной связи по состоянию для системы:
+              \begin{displaymath}
+                \dot x = (A+B_u\Theta)x + B_v v, x(0)=0
+              \end{displaymath}
+              Выходы системы описываются выражениями:
+              \begin{displaymath}
+                z_k=(C_k+B_u\Theta),k=\overline{1,N}
+              \end{displaymath}
+              Критерии оптимальности:
+              \begin{displaymath}
+                J_k(\Theta)=\sup_{v\in L_2} \frac{\max_{1\leqslant i \leqslant n_k} \sup_{t\geqslant 0}|z_k^{(i)}(\Theta,t)|}{||v||_2} \rightarrow\min_{\Theta}
+              \end{displaymath}
+            Ограничение на устойчивость системы:
+            \begin{displaymath}
+              g_0(\Theta)=\min_{j}\re(\lambda_j(A+B_u\Theta)) < 0
+            \end{displaymath}
+            Многокритериальная задача сводится к скалярной методом уступок или с помощью свёртки Гермейера.
+            В [] доказано, что использование последней позволяет найти всё множество Парето.
+          \end{block}
+            \begin{block}{Задача глобальной оптимизации}
+              Постановка задачи с ограничениями:
+              \begin{displaymath}
+                \varphi(y^*)=\min\{\varphi(y):y\in D\}, D=\{x\in \mathbf{R}^n: g_j(x) \leqslant 0, j=\overline{1,m}\}
+              \end{displaymath}
+              Предполагается, что все функции задачи удовлетворяют условию Липшица:
+              \begin{displaymath}
+              |\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
+              \end{displaymath}
+            \end{block}
             \begin{block}{Метод глобальной оптимизмции}
-            \end{block}
-            \begin{block}{Задача поиска оптимального управления}
+              Метод Стронгина с редукцией размерности с помощью кривой Пеано.\
+              \begin{minipage}[t]{.48\textwidth}
+              \begin{figure}
+                  \centering
+                  \includegraphics[scale=1.9]{images/peano1.png}
+              \end{figure}
+            \end{minipage}
             \end{block}
@@
         \begin{column}[t]{0.48\paperwidth}
+          \begin{block}{Способ распараллеливания}
+
+          \end{block}
             \begin{block}{Результаты}

### 2017-03-08 2017/PAVT/poster/poster.tex
PAVT: improve method description, add references

@@
             Многокритериальная задача сводится к скалярной методом уступок или с помощью свёртки Гермейера.
-            В [] доказано, что использование последней позволяет найти всё множество Парето.
+            В \cite{optControl} доказано, что использование последней позволяет найти всё множество Парето.
           \end{block}
@@
               \begin{displaymath}
-              |\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
+              |f(y_1)-f(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
               \end{displaymath}
+
+              Целевая функция и ограничения могут быть невыпуклы, многоэкстремальны, недифференцируемы.
             \end{block}
             \begin{block}{Метод глобальной оптимизмции}
-              Метод Стронгина с редукцией размерности с помощью кривой Пеано.\
+              Для оптимизации используется одномерный метод Стронгина.
+              Редукция размерности осуществляется с помощью кривой Пеано.
+
+              Общая схема одной итерации одномерного метода:
+              \begin{itemize}
+                \justifying
+                \item Упорядочить точки предшествующих испытаний в порядке возрастания их координат: \(a=x_{0}<...<x_{i}<...<x_{k}=b\).
+                \item Вычислить для каждого интервала \((x_{i-1};x_{i}),1\leq i\leq k\)  характеристику \(R(i)\) .
+                \item Определить интервал \((x_{t-1};x_{t})\) , которому соответствует максимальная характеристика \(R(t)=\max\{R(i),1\leq i\leq k\}\).
+                \item Провести следующее испытание в точке \(x^{k+1}=d(t)\in (x_{t-1};x_{t})\) , где \(d(t)\)  — правило размещения точки следующего испытания в интервале с номером \(t\).
+                \item Проверить выполнение критерия остановки \(x_{t}-x_{t-1}<\varepsilon\).
+              \end{itemize}
+              Подробное описание метода можно найти в \cite{optParallelBook}.
+
               \begin{minipage}[t]{.48\textwidth}
@@
                   \centering
-                  \includegraphics[scale=1.9]{images/peano1.png}
+                  \includegraphics[scale=1.27]{images/peano.png}
               \end{figure}
               \end{minipage}
+
             \end{block}
@@
             \end{block}
+            \begin{block}{Ссылки}
+              \printbibliography
+            \end{block}
         \end{column}

### 2017-03-26 2017/PAVT/poster/poster.tex
PAVT: improve parallel method description, add a draft of the results section

@@
             В \cite{optControl} доказано, что использование последней позволяет найти всё множество Парето.
+
+            Под данную постановку подходят задачи виброизоляции. Объект защиты представлен многомассовой механической системой, состоящей из
+            \(n\) материальных точек, связанных одинаковыми линейными упруго-диссипативными элементами между собой и основанием.
+
+            На данном этапе рассматривается задача виброизоляции системы из 10 точек. Поскольку на практике наблюдать состояние системы целиком
+            затратно, размерность пространства параметров была выбрана равной 3 (не все элементы вектора-состояния участвуют в формировании обратной связи).
           \end{block}
@@
               Общая схема одной итерации одномерного метода:
-              \begin{itemize}
+              \begin{enumerate}
                 \justifying
@@
                 \item Проверить выполнение критерия остановки \(x_{t}-x_{t-1}<\varepsilon\).
-              \end{itemize}
-              Подробное описание метода можно найти в \cite{optParallelBook}.
+              \end{enumerate}
+              Подробное описание метода и индексной схемы учёта ограничений можно найти в \cite{strOptBook}.
 
@@
         \begin{column}[t]{0.48\paperwidth}
-          \begin{block}{Способ распараллеливания}
-
+          \begin{block}{Параллельные версии метода оптимизации}
+            Существует несколько способов распареллеливания алгоритма глобального поиска (АГП):
+            \begin{itemize}
+              \justifying
+              \item Распараллеливание по характеристикам в рамках системы с общей памятью.
+              На шаге 2 вместо одного интервала выбираются \(p\) интервалов с наилучшими характеристиками и в них параллельно проводятся испытания.
+              \item Распараллеливание по развёрткам в рамках системы с раздельной памятью. На каждом узле системы работает копия метода, использующая
+              уникальную развёртку. Копии метода обмениваются многомерными точками, однако одномерные прообразы этих точек различны для каждого метода.
+              При использовании \(L\) развёрток каждый метод дополнительно получает \(L-1\) точку в свою поисковую информацию на каждой итерации, что
+              ускоряет его сходимость.
+              \item Сочетание указанных выше подходов.
+            \end{itemize}
+          В \cite{optParallelBook} перечисленные схемы описаны более подробно.
           \end{block}
           \begin{block}{Результаты}
+            В таблице приведены результаты применения распараллеливания по характеристикам, а также
+            результаты, полученные при использовании комбинированного подхода с двумя развёртками.
+            Поскольку размерность задачи оптимизации невелика, добавление дополнительных развёрток не привело к ускорению.
+
+            Достигнутым практическим результатом является построение парето границы на плоскости критериев (рис. \ref{fig:pareto}).
+            \begin{figure}
+                \label{fig:pareto}
+                \centering
+                \includegraphics[scale=1.2]{images/solution.pdf}
+                \caption{Парето-граница на плоскости критериев}
+            \end{figure}
           \end{block}
-            \begin{block}{Ссылки}
+          \begin{block}{Дальнейшая работа}
+            В процессе решения задачи выяснилось, что критерии обладают большой константой Липшица вблизи границы устойчивости системы, что затрудняет оптимизацию и
+            приводит к нестабильности результатов параллельных методов. Для решения этой проблемы требуется изменение решающих првил одномерного метода.
+          \end{block}
+          \begin{block}{Литература}
             \printbibliography

### 2017-04-01 2017/PAVT/poster/poster.tex
PAVT: add draft of performance table
PAVT: update experiments table

@@
             \end{displaymath}
-            Многокритериальная задача сводится к скалярной методом уступок или с помощью свёртки Гермейера.
+            Многокритериальная задача сводится к скалярной методом \(\varepsilon\)-ограничений или с помощью свёртки Гермейера.
             В \cite{optControl} доказано, что использование последней позволяет найти всё множество Парето.
@@
           \begin{block}{Результаты}
-            В таблице приведены результаты применения распараллеливания по характеристикам, а также
+            В таблице \ref{tab:parallelResults} приведены результаты применения распараллеливания по характеристикам, а также
             результаты, полученные при использовании комбинированного подхода с двумя развёртками.
             Поскольку размерность задачи оптимизации невелика, добавление дополнительных развёрток не привело к ускорению.
+
+            \begin{table}[h]
+            \caption{Результаты параллельных запусков на двух узлах кластера}
+            \label{tab:parallelResults}
+            \begin{tabular}{c|cccc}
+              $L$ & $p=1$  & $p=2$ & $p=4$ & $p=8$\\
+              \hline
+              1  &  75,09 (1) & 59,58 (1,26) & 39.60 (1,89) & 12.28 (6,11) \\
+              2  &  58,55 (1,28) & 46,14 (1,62) & 18.09 (4,15) & 9.21 (8,15) \\
+              \hline
+            \end{tabular}
+            \end{table}
 

### 2017-04-11 2017/EUROGEN/eurogen.tex
eurogen: add a placeholder for paper

% EUROGEN 2015 - Template for the extended abstract
%defining format issues
%**********************

%Document Settings___________________________________________________________________

\documentclass[10pt,twocolumn,a4paper]{article}
\special{papersize=210mm,297mm}

%\makeatletter
%\renewcommand\section{\@startsection{section}{1}{\z@}%
%                                  {-3.5ex \@plus -1ex \@minus -.2ex}%
%                                  {2.3ex \@plus.2ex}%
%                                  {\normalfont\normalsize\bfseries}}
%\makeatother

%\makeatletter
%\renewcommand\subsection{\@startsection{subsection}{1}{\z@}%
%                                  {-3.5ex \@plus -1ex \@minus -.2ex}%
%                                  {2.3ex \@plus.2ex}%
%                                  {\normalfont\normalsize\em}}
%\makeatother

%________________________________________________________________________________

%Miscellanneous__________________________________________________________________

\sloppy % better line breaks
\newcommand{\authorname}{\textbf}
\newcommand{\authorgroup}{\emph}
\newcommand{\authoraddress}{\emph}
\newcommand{\authormail}{\emph}
%________________________________________________________________________________

%Packages________________________________________________________________________

%\usepackage[latin1]{inputenc}
\usepackage[utf8]{inputenx}
\usepackage[compact]{titlesec}
%\usepackage[pdftex]{graphicx}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,mathrsfs}
\usepackage{fancyhdr}
\usepackage{calc}
\usepackage{titling}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage[superscript]{cite}
\usepackage{titlesec}
\usepackage[none]{hyphenat}
%________________________________________________________________________________
\titleformat*{\section}{\normalfont\normalsize\bfseries}
\titleformat*{\subsection}{\normalfont\normalsize\em}
\voffset = -1cm
\textheight = 23.7cm
\hoffset = -0.5cm
\textwidth = 17cm
\setlength\parindent{0.5cm}
%Headers_________________________________________________________________________

\pagestyle{fancy}
\fancyhead[L]{EUROGEN 2017}\fancyhead[R]{September 13-15, 2017, Madrid, Spain}
\fancyheadoffset[LE,RO]{\marginparsep + \marginparwidth}
\setlength{\topmargin}{0pt}
%________________________________________________________________________________

%Maketitle Conf__________________________________________________________________
\pretitle{\begin{center}\Large}
\posttitle{\end{center}}

\preauthor{\begin{center}\normalsize \lineskip 0.5em\begin{tabular}[t]{c}}
\postauthor{\end{tabular}\par\end{center}}

\predate{\begin{center}\tiny}
\postdate{\par\end{center}}
%________________________________________________________________________________
%--------------------------------------------------------------------------------

%Article data entry_____________________________________________________________

\title{\vspace{-5ex}\textbf{Instructions to Authors for the Preparation of Extended Abstract}}
\author{
  \authorname{First I. Author*}\\
  \authorgroup{International Center for Numerical Methods in Engineering (CIMNE)}\\
  \authoraddress{Universidad Politécnica de Cataluña,Campus Norte UPC, 08034 Barcelona, Spain}\\
  \authormail{Email: cimne@cimne.upc.edu}\\\\
  %--------------------------
  \authorname{Second B. Author}\\
  \authorgroup{Spanish Association for Numerical Methods in Engineering (SEMNI)}\\
  \authoraddress{Edificio C1, Campus Norte UPC, Gran Capitán s/n, 08034 Barcelona, Spain}\\
  \authormail{Email:semni@cimne.upc.edu}
  %}
}
\date{}

%_______________________________________________________________________________

%Article Body___________________________________________________________________

\begin{document}

%Abstract goes here--------------------------------------------------------------

\twocolumn[

\maketitle

\thispagestyle{fancy} %Insert header in title page

\begin{center}
\line(1,0){465}
\end{center}

\begin{@twocolumnfalse}

\textbf{Summary}\\

The purpose of this document is to provide an example of how authors must format their extended abstracts. The summary should be a very brief description of the work, which should enable an easy indexing. It should not exceed 200 words in length. It must be single spaced, even justified across the full width of the page.\\

\emph{\textbf{Keywords:} list up to 8 (eight) words or group of words in order of priority, separated by commas, which will be used for indexing purposes.}
\begin{center}
\line(1,0){465}
\end{center}

\end{@twocolumnfalse}

]

%--------------------------------------------------------------------------------

\section{Introduction}

Extended abstracts must be delivered through the web submission page, completely “camera-ready” (postscript file or PDF).

\section{General specifications}

The extended abstract must be written in English within a printing box of 17cmx23.7cm, centered in the page. The paper including figures, tables and reference mast have a minimum length of 4 pages and must not exceed 8 pages. Any processing systems can be used (Tex, Latex, Word….).

\section{Title, Authors, Affiliation}

\subsection{Title}

The title should be written centered, in 14pt, boldface Roman. It should be single spaced if the title is more than one line long.

\subsection{Author}

The author's name should include first name, middle initial and surname. It should be written centered, (as shown in the first page of these instructions) in 10pt boldface Roman. The presenting author should be marked with an asterisk.

\subsection{Affiliation}

Author's affiliation should be written centered, (as shown in the first page of these instructions) in 10pt italic Roman.

\section{Text}

The normal text should be written single-spaced, justified, using 10pt (Times New) Roman in two columns.

The first line of each paragraph must be indented 0.5cm (except the first paragraph after a heading). There is not inter-paragraph spacing.

\section{Format of references}

References should be quoted in the text by superscript numbers\cite{Baeck:92, Deb:2002}, and grouped together at the end of the paper in numerical order as shown in these instructions.

The entire list of references at the end of the paper should not be longer than 1/8 of the paper (i.e., 1 full column if the paper is 4 pages long, and 2 columns if the paper is 8 pages long).

\bibliography{bibliography}{}
\bibliographystyle{nature}

\end{document}
__________________________________________________________________________

### 2017-04-11 2017/EUROGEN/eurogen.tex
EUROGEN: update paper

@@
 \usepackage[none]{hyphenat}
+\usepackage{enumitem}
 %________________________________________________________________________________
@@
 
-\title{\vspace{-5ex}\textbf{Instructions to Authors for the Preparation of Extended Abstract}}
+\title{\vspace{-5ex}\textbf{Multipoint Approximation Method for Large Scale Design Optimization Problems Under Uncertainty}}
 \author{
@@
 
-Extended abstracts must be delivered through the web submission page, completely “camera-ready” (postscript file or PDF).
+In industrial design optimization problems, random inaccuracies of the production process may lead to the designer’s inability to fully control the design variables.
+Therefore, there can be a discrepancy between the products \textit{as designed} and \textit{as produced} that results in actual performance being different from the performance expected at the design stage.
+Possible discrepancy between the design and the manufactured product becomes even more important if the latter violates some critical constraints, e.g.
+safety regulations.
 
-\section{General specifications}
+In addition, the system’s performance may depend on parameters of random nature that are beyond the designer’s control, e.g.
+the operation conditions.
+Therefore,  accounting for both of these types of uncertainties  is crucial for achieving robust and reliable performance.
 
-The extended abstract must be written in English within a printing box of 17cmx23.7cm, centered in the page.
-The paper including figures, tables and reference mast have a minimum length of 4 pages and must not exceed 8 pages.
-Any processing systems can be used (Tex, Latex, Word….).
+Optimisation under uncertainty deals with responses $F(\pmb x,\pmb y)$ that depend on deterministic design variables $\pmb x$ and random ‘environmental’ variables $\pmb y$.
+Uncertainty in the design variables can also be modelled by introducing additional random variables, e.g.
+as additive or multiplicative noise.
+In this case, the responses $F(\pmb x,\pmb y)$ receive as input the sum $\pmb x+\pmb u$ (if the noise is additive), where $\pmb u$ is a random variable with a known probability distribution, yielding responses depending on $\pmb x,\; \pmb u$ and $\pmb y$.
+Randomness in $\pmb u$ and $\pmb y$ induces randomness in the responses $F(\pmb x+\pmb u,\pmb y)$.
+To convert them into deterministic functions, that can be optimised, a mapping $R$, called the \textit{risk measure}[2] or \textit{robustness measure}[3], is applied to the responses.
+The result is then a deterministic function of the design variables: $\widetilde{F}(\pmb x)=R(F(\pmb x+\pmb u, \pmb y))$.
+The choice of a particular risk measure $R$ reflects the designer's attitude towards the risk of sub-optimal performance or constraint violation.
+For example, $R$ can be a combination of the mean and standard deviation of $F(\pmb x+ \pmb u,\pmb y)$ with respect to the joint probability distribution of $\pmb u$ and $\pmb y$ or a quantile of the distribution of $F(\pmb x+ \pmb u,\pmb y)$, i.e.
+such a value that the random variable of $F(\pmb x+\pmb u,\pmb y)$ exceeds it with a given, typically small, probability.
 
-\section{Title, Authors, Affiliation}
+Computations of the risk measures typically involve evaluation of high-dimensional integrals, which becomes a challenging task if the number of design variables and/or environmental variables is large.
+The problem becomes even more difficult if the original responses are computationally expensive, which is the case, for example, in computational fluid dynamics (CFD), where one function evaluation may take hours of even days.
+Therefore, direct computation of the risk measures using original responses becomes infeasible and therefore approximations of responses, also referred to as metamodels, need to be used.
 
-\subsection{Title}
+In problems with a large (in the order of hundreds) number of design variables, the multipoint approximation method (MAM4,5,6) proved to be efficient, e.g.
+in turbomachinery applications7,8,9.
+This method is an iterative optimization technique based on mid-range approximations built in trust regions.
+A trust region is a sub-domain of the design space in which a set of design points, produced according to a small-scale design of experiments (DoE), are evaluated.
+These and a subset of previously evaluated design points are used to build metamodels of the objective and constraint functions that are considered to be valid within a current trust region.
+The trust region will then translate and change size as optimization progresses.
+The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values10,11 and occasional simulation failures12.
+The mid-range approximations used in the trust regions, as originally suggested in Ref.
+4 for structural optimization problems, are intrinsically linear functions (i.e.
+nonlinear functions that can be led to a linear form by a simple transformation) for individual sub-structures, and assembly of them for the whole structure.
+This was enhanced by the use of gradient-assisted metamodels,6 use of simplified numerical models that is also termed a multi-fidelity approach,13 and the use of analytical models derived by Genetic Programming14.
+One of the recent developments15 involved the use of approximation assemblies, i.e.
+a two stage approximation building process that is conceptually similar to the original one used in Ref.
+4 but is free from the limitation that lower level approximations are linked to individual substructures.
 
-The title should be written centered, in 14pt, boldface Roman. It should be single spaced if the title is more than one line long.
+This paper presents a new development in the Multipoint Approximation Method that makes it capable of handling problems with uncertainty in design variables as well as in additional ‘environmental’ variables.
+The approach relies on approximations built in the combined space of design variables and environmental variables, and subsequent application of a risk measure and optimization with respect to the deterministic design variables, all within the iterative trust-region-based framework of MAM.
 
-\subsection{Author}
+\section{The Multipoint Approximation Method}
 
-The author's name should include first name, middle initial and surname.
-It should be written centered, (as shown in the first page of these instructions) in 10pt boldface Roman.
-The presenting author should be marked with an asterisk.
+It is useful to start with a brief description of the deterministic version of MAM. A typical formulation of a constrained optimization problem that MAM works with is as follows:
+\begin{equation}
+  \label{eq:problem}
+  \begin{array}{c}
+  \min_{a_i \le x_i \le b_i}F_0(\pmb x) \\
+  s.t.\; F_j(\pmb x) \le 1,\; j=1,\dots ,M,
+  \end{array}
+\end{equation}
+where $\pmb x$ is the vector of design variables, $\pmb a$ and $\pmb b$ are the lower and upper bounds for the design variables, respectively, $F_0(\pmb x)$ is the objective function, and $F_j(\pmb x)$ are the constraints.
+The numbers of design variables and constraints are $n$ and $M$, respectively.
+MAM attempts to solve this problem by using approximations of the objective function and constraints in a series of trust regions.
+The trust region strategy seeks to zoom in on the region where the constrained minimum is achieved.
+It aims at finding a trust region that is sufficiently small for the approximations to be of sufficiently good quality to improve the design, and that contains the point of the constrained minimum as its interior point.
+The main loop of the MAM is organized as follows.
 
-\subsection{Affiliation}
+Algorithm 1 (deterministic MAM).
+\begin{enumerate}
+\item Initialization: choose a starting point $\pmb x^0$ and initial trust region $[\pmb a^0, \pmb b^0]$ such that $\pmb x^0 \in [\pmb a^0, \pmb b^0]$.
+\item At the $k$-th iteration the current approximation to the constrained minimum is $\pmb x^k$, the current trust region is $[\pmb a^k, \pmb b^k] \subset [\pmb a^0, \pmb b^0]$.
+  \begin{enumerate}[label=(\alph*)]
+    \item Design of Experiments (DoE): a set of points $\pmb x_k^i \in [\pmb a^k, \pmb b^k]$ is chosen to be used for approximation building.
+Responses are evaluated at the DoE points and approximations are built using the obtained values.
+Currently, the pool of approximation methods available in MAM consists of a metamodel assemblies15 and the moving least-squares metamodels16-19.
+Other metamodel types could be used as well.
 
-Author's affiliation should be written centered, (as shown in the first page of these instructions) in 10pt italic Roman.
+    Denote the approximate objective function and constraints by $\widetilde{F}_0(\pmb x)$ and $\widetilde{F}_j(\pmb x)$, respectively.
+    \item The original optimization problem (\ref{eq:problem}) is replaced by the following problem:
+    \begin{equation}
+      \label{eq:problem_approx}
+      \begin{array}{c}
+      \min_{a_i^k \le x_i ^k\le b_i}\widetilde{F}_0(\pmb x) \\
+      s.t.\; \widetilde{F}_j(\pmb x) \le 1,\; j=1,\dots ,M,
+      \end{array}
+    \end{equation}
+    The approximate problem (\ref{eq:problem_approx}) is solved using Sequential Quadratic Programming (SQP) and the solution of this problem   determines the centre of the next trust region.
+    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region10.
+    \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $\pmb x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations).
+If the termination criterion is satisfied, the algorithm proceeds to step 3.
+Otherwise, it returns to step 2.
+  \end{enumerate}
+  \item Optimization terminates. The obtained approximation to the solution of the problem (\ref{eq:problem}) is $\pmb x^{k+1}$.
+\end{enumerate}
 
-\section{Text}
+\section{Optimization Under Uncertainty}
 
-The normal text should be written single-spaced, justified, using 10pt (Times New) Roman in two columns.
+The presence of uncertainties in the responses requires certain reformulations in the problem statement (\ref{eq:problem}).
+Informally, the optimization problem with uncertainty can be written as follows:
+\begin{equation}
+  \label{eq:problem_unc}
+  \begin{array}{c}
+  \min_{a_i \le x_i \le b_i}F_0(\pmb x + \pmb u, \pmb y) \\
+  s.t.\; F_j(\pmb x + \pmb u, \pmb y) \le 1,\; j=1,\dots ,M,
+  \end{array}
+\end{equation}
+where the random variable $\pmb u$ represents (additive) noise in the design variables and $\pmb y$ is the environmental variable, representing input to the responses that cannot be influenced by the designer.
+This problem statement is not precise since the random responses $F_0 (\pmb x+\pmb u,\pmb y),\; F_j (\pmb x+\pmb u,\pmb y)$ cannot be directly optimized.
+Therefore, a risk measure $R$ needs to be applied to the objective and the constraints:
+\begin{equation}
+  \label{eq:problem_unc_r}
+  \begin{array}{c}
+  \min_{a_i \le x_i \le b_i}R_{u,y}(F_0)(\pmb x) \\
+  s.t.\; R_{u,y}(F_j)(\pmb x) \le 1,\; j=1,\dots ,M,
+  \end{array}
+\end{equation}
+where the subscript in $R_{\pmb u,\pmb y}$ indicates that the variables $\pmb u$ and $\pmb y$ ‘collapse’ when the risk measure is applied and the result only depends on $\pmb x$.
+The choices of risk measure will be discussed in Section \ref{sec:risk}.
+Here the changes in the Algorithm 1 that are necessary to solve the problem (\ref{eq:problem_unc_r}) are discussed.
 
-The first line of each paragraph must be indented 0.5cm (except the first paragraph after a heading). There is not inter-paragraph spacing.
+First, it needs to be decided on whether to build metamodels for the risk measure $R_{u,y}(F_j)(\pmb x)$ or for the original responses $F_j (\pmb x+ \pmb u,\pmb y)$.
+The first option may seem attractive due to the simplicity of its implementation in Algorithm 1 but the deciding argument has to be that of the computational cost associated with the evaluation of either function.
+On the one hand, one evaluation of the function $R_{u,y}(F_j)(\pmb x)$ involves computing one or several integrals of the original response $F_j(\pmb x+ \pmb u,\pmb y)$ over the distribution of $\pmb u$ and $\pmb y$, which requires many evaluations of the original response.
+On the other hand, the dimensionality of the argument of $R_{u,y}(F_j)(\pmb x)$ is less than that of the argument of $F_j(\pmb x+ \pmb u,\pmb y)$, therefore, in principle, fewer function evaluations are needed to approximate the former than the latter.
+In practice, this may be a matter of a trade-off between the two abovementioned effects that depends on the proportion between the numbers of design variables $n$ and the environmental variables $n_e$ and on the complexity of the integration method and metamodeling method used.
+If the number if design variables is significantly larger than the number of environmental variables, which is often the case and which is assumed in this paper, building metamodels for the original responses $F_j(\pmb x+ \pmb u,\pmb y)$ is preferred.
 
-\section{Format of references}
+This choice requires some changes in the MAM algorithm, since metamodels are now built in a larger space than the one where optimisation is performed in. The changes are summarised below.
 
-References should be quoted in the text by superscript numbers\cite{Baeck:92, Deb:2002}, and grouped together at the end of the paper in numerical order as shown in these instructions.
+Algorithm 2 (MAM under uncertainty).
+\begin{enumerate}
+  \item Initialization: choose a starting point $\pmb x_0 \in {\Bbb R}^n$.
+Choose an initial trust region $[\pmb a^0,\pmb b^0]$ in the combined space of design variables and environmental variables.
+Ensure that the initial guess $\pmb x_0$ belongs to the projection of the initial trust region onto the subspace of design variables.
+The projection of the initial trust region onto the subspace of environmental variables must cover the region where the realisations of the random variable y are likely to be.
+The coordinates of the trust region corresponding to the environmental variables will not be changed in the course of optimisation.
+Typically, the trust region along the coordinates corresponding to the $i$-th environmental variable is centred at the mean of the $i$-th environmental variable and has the width of six of its standard deviations.
+  \item At the $k$-th iteration the current approximation to the constrained minimum is $\pmb x^k$, the current trust region is $[\pmb a^k,\pmb b^k]$.
+  \begin{enumerate}[label=(\alph*)]
+    \item Design of Experiments (DoE): a set of points $(x_i^k,y_i^k)\in [\pmb a^k,\pmb b^k]$ is chosen that will be used to build approximations.
+Responses are evaluated at the DoE points and approximations are built using the obtained values.
+    Denote the approximate objective function and constraints by $\widetilde{F}_0(\pmb x, \pmb y)$ and $\widetilde{F}_j(\pmb x, \pmb y)$, respectively.
+    \item The original optimization problem is replaced by the following problem:
+    \begin{equation}
+      \label{eq:problem_u_tilde}
+      \begin{array}{c}
+      \min_{a_i \le x_i \le b_i}R_{u,y}(\widetilde{F}_0)(\pmb x) \\
+      s.t.\; R_{u,y}(\widetilde{F}_j)(\pmb x) \le 1,\; j=1,\dots ,M,
+      \end{array}
+    \end{equation}
+    The risk measures are calculated using the metamodels obtained in the previous step.
+Integrals that arise during this process are calculated with respect to the joint distribution of $\pmb u$ and $\pmb y$.
+This distribution is assumed to be fixed.
+    The approximate problem (\ref{eq:problem_u_tilde}) is solved using Sequential Quadratic Programming (SQP) and the centre of the next trust region $\pmb x^{k+1}$ is determined as the solution of this problem.
+The optimisation runs in a subspace of the space where metamodels are built: metamodels are defined in the combined space of design variables and environmental variables, whilst optimisation runs in the subspace of the design variables.
+    \item The trust region is updated.
+The size of the trust region along the dimensions corresponding to the environmental variables is unchanged.
+The size of the next trust region along the dimensions corresponding to the design variables is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region according to the trust region strategy explained in Ref.
+10.
+    \item The termination criterion is checked. If the termination criterion is satisfied, the algorithm proceeds to step 3. Otherwise, it returns to step 2
+  \end{enumerate}
+  \item Optimization terminates. The obtained approximation to the optimum is $\pmb x^{k+1}$.
+\end{enumerate}
 
-The entire list of references at the end of the paper should not be longer than 1/8 of the paper (i.e., 1 full column if the paper is 4 pages long, and 2 columns if the paper is 8 pages long).
+Summarising, the key changes introduced to account for uncertainties are (i) metamodels built in the combined space of design variables and environmental variables; (ii) optimisation performed in a subspace of the space where metamodels are defined; (iii) additive noise in the design variables; (iv) application of risk measures to transform random responses into deterministic functions.
+
+\section{Risk Measures}
+\label{sec:risk}
+
+As it has been already said, a risk measure is a mapping that transforms a random response $F(\pmb x+ \pmb u,\pmb y)$  into a deterministic function that can be optimized.
+Next we provide some examples of risk measures2,3.
+
+\begin{itemize}
+  \item $R(F)=\mu(F)$
+\end{itemize}
+
+\section{Conclusion}
+
+Recent developments in the Multipoint Approximation Method made it capable of solving large scale problems with uncertainty in parameters modelled as additional (‘environmental’) random variables and uncertainty in the design variables modelled as additive noise.
+The approach taken combines metamodels built in the combined space of design variables and environmental variables, application of risk measures that map random responses to deterministic functions, and optimisation in a subspace of the space where metamodels are defined, all within the iterative and trust-region-based framework of MAM.
+Performance is demonstrated on a benchmark example of structural optimisation known as the scalable cantilevered beam.
+
+\section*{Acknowledgement}
+The authors are grateful for the support provided by the Russian Science Foundation, project No. 16-11-10150.
 

### 2017-04-12 2017/EUROGEN/eurogen.tex
EUROGEN: update sections, add a picture

@@
 \begin{itemize}
-  \item $R(F)=\mu(F)$
+  \item $R(F)=\mu(F)$ where $μ(F)$ is the mean of $F$.
+This choice corresponds to optimising average performance of the system.
+In typical engineering applications, however, this is not sufficient and more conservative risk measures need to be chosen.
+  \item $R(F)=\sup(F)$.
+This is the most conservative case, in which one is concerned with the worst case performance of the system.
+The expression $\sup(F)$ stands for the least of all values that are greater than the random response $F$ with probability 1.
+If the support of the probability density function (PDF) of $F$ is not compact (i.e.
+$F$ can take arbitrary large values with non-zero probability) than $\sup(F)$ will be infinite.
+  \item $R(F)=\mu(F)+k\sigma (F)$, where $\sigma(F)$ is the standard deviation of $F$ and $k>0$ is a constant.
+This risk measure is sometimes referred to as the safety margin.
+Optimisation using this risk measure is referred to as robust design optimisation22,23,24 and reflects the paradigms of a ‘$3\sigmaσ$-design’, a ‘$6\sigmaσ$-design’, etc.
+  \item $R(F)=q_F(\alpha)$, where $q_F(\alpha)$ is the value of the quantile function, or simply the quantile, corresponding to a given probability $\alpha\in (0,1)$.
+This choice corresponds to the paradigm of reliability-based design optimization25,23,26.
+In this approach, one wants to keep the probability of the system failing (i.e.
+violating the constraint or underperforming, depending on the context) below a given threshold.
+The value $1−\alpha$ is the desired probability of failure.
+Although rather intuitive, this choice has known computational issues which are discussed by Rockafellar and Royset27.
 \end{itemize}
+
+For this study, we chose as risk measure the safety margin $R(F)=\mu(F)+k\sigma(F)$, aiming at covering both robust design optimisation and reliability design optimisation.
+Let us briefly discuss its computation.
+
+To compute the safety margin risk measure, one needs to compute two integrals:
+\begin{displaymath}
+  \mu (F)=E[F]=
+\end{displaymath}
+
+The dimension of these integrals is $n+n_e$.
+To compute them efficiently, methods of high-dimensional integration must be employed.
+Here, quasi-Monte-Carlo integration30 based on Niederreiter sequences31 is used.
+
+\section{Numerical Example}
+
+The example considered in this study is a classical engineering optimisation problem known as the scalable cantilevered beam34.
+The engineering object to be optimised is shown in Fig.
+2 taken from Ref.
+34.
+
+%picture
+
+The design variables are the widths $b_i$ and heights $h_i$ of the segments.
+The number of segments $N$ can be chosen abitrary.
+The total length of the beam is 500 cm, the lenghths of the segments are $l_i=500/N$  cm.
+There are $N$ geometric constraints (the aspect ratios of each block (i.e.
+heights divided by widths) should not exceed 20) and $N$ constraints on the stress, calculated at the left end of each segment (stresses should not exceed  $\sigma=14000 N/cm^2$).
+There is also also a constraint on the displacement at the tip, which should not exceed 2.5 cm.
+The load is $P = 50 000 N$, the Young's modulus is $E=2\cdot 10^7  N/cm^2$.
+
+The deflection $y_i$ at the right hand of the $i$-th segment is given by the following recursive formulas:
+\begin{displaymath}
+  \mu (F)=E[F]=
+\end{displaymath}
+The moment of inertia of a segment is $I_i=\frac{b_i h_i^3}{12}$ and the bending moment at the left end is $M_i=P[L+l_i- \sum_{j=1}^{i}  l_j ]$.
+The maximum bending stress in the segment $i$ is then given by the following formula:
+\begin{displaymath}
+  \mu (F)=E[F]=
+\end{displaymath}
+
+We are looking for a design with smallest volume $V = \sum_{i=1}^N b_i h_i l_i$.
+The widths $b_i$ vary from 1.0 to 10.0 cm and the heights $h_i$ from 5.0 to 100.0 cm.
+The optimisation problem is formulated as follows:
+\begin{displaymath}
+  \mu (F)=E[F]=
+\end{displaymath}
+
+With $N=50$ segments (corresponding to 100 design variables) the SQP solution of this problem is $V = 63704.598 cm^3$.
+MAM obtained the solution $V = 63935.360 cm^3$, using 1500 function evaluations (as compared to almost 8000 evaluations used by SQP).
+The number of points in a trust region used to build approximations was 150.
+The solution obtained by MAM is very close to the reference solution obtained by SQP, except for the last design variable (the height of the last segment), which indicates that the problem is insensitive to this variable near the optimum, making it hard for the metamodels to capture this dependence.
+Both SQP and MAM solutions are, however, feasible and differ only slightly in the value of the objective function.
+
+Now let us add uncertainty to the problem, adding Gaussian noise with zero mean and variance 0.1 to each design variable.
+Using the μ(F) + kσ(F) risk measure, we obtain the following problem statement:
+\begin{displaymath}
+  \mu (F)=E[F]=
+\end{displaymath}
+
+The notation $\sigma$ was used to denote the variance of a response, but it also stands for the maximum bending stresses $\sigma_i$, resulting in somewhat confusing notation $\sigma(\sigma_i )$.
+Hopefully, this explanation eliminates the confusion.
+Optimising $\mu(F) + k\sigma(F)$ with $k = 3.0$ using SQP on original functions, we obtain the value of the deterministic objective (volume of the structure) at the robust optimum $V = 67598.049 cm^3$.
+To compute $\mu(F)$ and $\sigma(F)$ at each iteration, 1000 points were used.
+Multiplied by the number of evaluations of the robust objective $\mu(F) + k\sigma(F)$ (approximately 8000), we get a rough estimate of $8\cdot 10^6$ of the total number of points that SQP used to find this solution.
+
+Solving the same problem using the new version of MAM suitable for problems with uncertainty, we obtain the value of the deterministic objective (volume of the structure) at the robust optimum $V = 67799.335 cm^3$.
+To obtain this result, MAM again used 150 points per trust region to build metamodels, which resulted in a total of 1500 function evaluations, which is orders of magnitude smaller than the number of points required if no metamodels are used.
+
+The $\mu(F ) + k\sigma(F)$ risk measure uses quantities that are estimated in the middle of the distribution --- the mean and the standard deviation.
+On the one hand, this has the advantage that samples drawn directly from the distribution of the random variables $\pmb u$ and $\pmb y$ are well suited to approximately compute this risk measure.
+On the other hand, probabilistic interpretation of the obtained results in this case is only possible under the normality assumption for the distribution of $F$; if the actual distribution of $F$ has heavier tails than the normal one, the actual probability of failure will be higher than the one  the system was designed for.
 

### 2017-04-12 2017/EUROGEN/eurogen.tex
EUROGEN: update chapters and references

@@
 \usepackage{enumitem}
+\usepackage{cite}
 %________________________________________________________________________________
@@
 \author{
-  \authorname{First I. Author*}\\
-  \authorgroup{International Center for Numerical Methods in Engineering (CIMNE)}\\
-  \authoraddress{Universidad Politécnica de Cataluña,Campus Norte UPC, 08034 Barcelona, Spain}\\
-  \authormail{Email: cimne@cimne.upc.edu}\\\\
+  \authorname{Vassili V. Toropov*}\\
+  \authorgroup{Professor of Aerospace Engineering, School of Engineering and Materials Science}\\
+  \authoraddress{Queen Mary University of London, London, UK}\\
+  \authormail{Email: v.v.toropov@qmul.ac.uk}\\\\
+  \authorgroup{Senior Research Fellow, Institute of Information Technology, Mathematics and Mechanics}\\
+  \authoraddress{Lobachevsky University of Nizhny Novgorod, Russia}\\\\
   %--------------------------
-  \authorname{Second B. Author}\\
-  \authorgroup{Spanish Association for Numerical Methods in Engineering (SEMNI)}\\
-  \authoraddress{Edificio C1, Campus Norte UPC, Gran Capitán s/n, 08034 Barcelona, Spain}\\
-  \authormail{Email:semni@cimne.upc.edu}
+  \authorname{Yury M. Korolev}\\
+  \authorgroup{Research Fellow, Institute of Mathematics and Image Computing}\\
+  \authoraddress{University of Lübeck, Germany}\\
+  \authormail{Email: yury.korolev@mic.uni-luebeck.de}\\\\
+  %--------------------------
+  \authorname{Victor P. Gergel, Konstantin A. Barkalov}\\
+  \authorgroup{Institute of Information Technology, Mathematics and Mechanics}\\
+  \authoraddress{Lobachevsky University of Nizhny Novgorod, Russia}
   %}
@@
 
-The purpose of this document is to provide an example of how authors must format their extended abstracts.
-The summary should be a very brief description of the work, which should enable an easy indexing.
-It should not exceed 200 words in length.
-It must be single spaced, even justified across the full width of the page.\\
+TBD\\
 
-\emph{\textbf{Keywords:} list up to 8 (eight) words or group of words in order of priority, separated by commas, which will be used for indexing purposes.}
+\emph{\textbf{Keywords:} TBD}
 \begin{center}
@@
 \section{Introduction}
-
+\label{sec:intro}
 In industrial design optimization problems, random inaccuracies of the production process may lead to the designer’s inability to fully control the design variables.
@@
 Randomness in $\pmb u$ and $\pmb y$ induces randomness in the responses $F(\pmb x+\pmb u,\pmb y)$.
-To convert them into deterministic functions, that can be optimised, a mapping $R$, called the \textit{risk measure}[2] or \textit{robustness measure}[3], is applied to the responses.
+To convert them into deterministic functions, that can be optimised, a mapping $R$, called the \textit{risk measure}\cite{RockafellarUryasev2000} or \textit{robustness measure}\cite{SchillingsSchulz2015}, is applied to the responses.
 The result is then a deterministic function of the design variables: $\widetilde{F}(\pmb x)=R(F(\pmb x+\pmb u, \pmb y))$.
@@
 
-In problems with a large (in the order of hundreds) number of design variables, the multipoint approximation method (MAM4,5,6) proved to be efficient, e.g.
-in turbomachinery applications7,8,9.
+In problems with a large (in the order of hundreds) number of design variables, the multipoint approximation method (MAM\cite{Toropov1989,Toropov1992,ToropovFilatov1993}) proved to be efficient, e.g.
+in turbomachinery applications\cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010}.
 This method is an iterative optimization technique based on mid-range approximations built in trust regions.
@@
 The trust region will then translate and change size as optimization progresses.
-The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values10,11 and occasional simulation failures12.
+The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values\cite{KeulenToropovMarkine1996,ToropovKeulenMarkine1996} and occasional simulation failures\cite{ToropovMarkineHolden1999}.
 The mid-range approximations used in the trust regions, as originally suggested in Ref.
-4 for structural optimization problems, are intrinsically linear functions (i.e.
+\cite{Toropov1989} for structural optimization problems, are intrinsically linear functions (i.e.
 nonlinear functions that can be led to a linear form by a simple transformation) for individual sub-structures, and assembly of them for the whole structure.
-This was enhanced by the use of gradient-assisted metamodels,6 use of simplified numerical models that is also termed a multi-fidelity approach,13 and the use of analytical models derived by Genetic Programming14.
-One of the recent developments15 involved the use of approximation assemblies, i.e.
+This was enhanced by the use of gradient-assisted metamodels,\cite{ToropovFilatov1993} use of simplified numerical models that is also termed a multi-fidelity approach,\cite{ToropovMarkine1996} and the use of analytical models derived by Genetic Programming\cite{ToropovAlvarez1998}.
+One of the recent developments\cite{PolynkinToropov2012} involved the use of approximation assemblies, i.e.
 a two stage approximation building process that is conceptually similar to the original one used in Ref.
-4 but is free from the limitation that lower level approximations are linked to individual substructures.
+\cite{Toropov1989} but is free from the limitation that lower level approximations are linked to individual substructures.
 
@@
 \section{The Multipoint Approximation Method}
+\label{sec:MAM}
 
@@
   \begin{array}{c}
-  \min_{a_i \le x_i \le b_i}F_0(\pmb x) \\
+  \min\limits_{a_i \le x_i \le b_i}F_0(\pmb x) \\
   s.t.\; F_j(\pmb x) \le 1,\; j=1,\dots ,M,
@@
       \begin{array}{c}
-      \min_{a_i^k \le x_i ^k\le b_i}\widetilde{F}_0(\pmb x) \\
+      \min\limits_{a_i^k \le x_i ^k\le b_i}\widetilde{F}_0(\pmb x) \\
       s.t.\; \widetilde{F}_j(\pmb x) \le 1,\; j=1,\dots ,M,
@@
   \begin{array}{c}
-  \min_{a_i \le x_i \le b_i}F_0(\pmb x + \pmb u, \pmb y) \\
+  \min\limits_{a_i \le x_i \le b_i}F_0(\pmb x + \pmb u, \pmb y) \\
   s.t.\; F_j(\pmb x + \pmb u, \pmb y) \le 1,\; j=1,\dots ,M,
@@
   \begin{array}{c}
-  \min_{a_i \le x_i \le b_i}R_{u,y}(F_0)(\pmb x) \\
-  s.t.\; R_{u,y}(F_j)(\pmb x) \le 1,\; j=1,\dots ,M,
+  \min\limits_{a_i \le x_i \le b_i}R_{\pmb u,\pmb y}(F_0)(\pmb x) \\
+  s.t.\; R_{\pmb u,\pmb y}(F_j)(\pmb x) \le 1,\; j=1,\dots ,M,
   \end{array}
@@
 
-First, it needs to be decided on whether to build metamodels for the risk measure $R_{u,y}(F_j)(\pmb x)$ or for the original responses $F_j (\pmb x+ \pmb u,\pmb y)$.
+First, it needs to be decided on whether to build metamodels for the risk measure $R_{\pmb u,\pmb y}(F_j)(\pmb x)$ or for the original responses $F_j (\pmb x+ \pmb u,\pmb y)$.
 The first option may seem attractive due to the simplicity of its implementation in Algorithm 1 but the deciding argument has to be that of the computational cost associated with the evaluation of either function.
-On the one hand, one evaluation of the function $R_{u,y}(F_j)(\pmb x)$ involves computing one or several integrals of the original response $F_j(\pmb x+ \pmb u,\pmb y)$ over the distribution of $\pmb u$ and $\pmb y$, which requires many evaluations of the original response.
-On the other hand, the dimensionality of the argument of $R_{u,y}(F_j)(\pmb x)$ is less than that of the argument of $F_j(\pmb x+ \pmb u,\pmb y)$, therefore, in principle, fewer function evaluations are needed to approximate the former than the latter.
+On the one hand, one evaluation of the function $R_{\pmb u,\pmb y}(F_j)(\pmb x)$ involves computing one or several integrals of the original response $F_j(\pmb x+ \pmb u,\pmb y)$ over the distribution of $\pmb u$ and $\pmb y$, which requires many evaluations of the original response.
+On the other hand, the dimensionality of the argument of $R_{\pmb u,\pmb y}(F_j)(\pmb x)$ is less than that of the argument of $F_j(\pmb x+ \pmb u,\pmb y)$, therefore, in principle, fewer function evaluations are needed to approximate the former than the latter.
 In practice, this may be a matter of a trade-off between the two abovementioned effects that depends on the proportion between the numbers of design variables $n$ and the environmental variables $n_e$ and on the complexity of the integration method and metamodeling method used.
@@
       \begin{array}{c}
-      \min_{a_i \le x_i \le b_i}R_{u,y}(\widetilde{F}_0)(\pmb x) \\
-      s.t.\; R_{u,y}(\widetilde{F}_j)(\pmb x) \le 1,\; j=1,\dots ,M,
+      \min\limits_{a_i \le x_i \le b_i}R_{\pmb u,\pmb y}(\widetilde{F}_0)(\pmb x) \\
+      s.t.\; R_{\pmb u,\pmb y}(\widetilde{F}_j)(\pmb x) \le 1,\; j=1,\dots ,M,
       \end{array}
@@
 The size of the next trust region along the dimensions corresponding to the design variables is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region according to the trust region strategy explained in Ref.
-10.
+\cite{KeulenToropovMarkine1996}.
     \item The termination criterion is checked. If the termination criterion is satisfied, the algorithm proceeds to step 3. Otherwise, it returns to step 2
@@
 \begin{itemize}
-  \item $R(F)=\mu(F)$ where $μ(F)$ is the mean of $F$.
+  \item $R(F)=\mu(F)$ where $\mu(F)$ is the mean of $F$.
 This choice corresponds to optimising average performance of the system.
@@
 $F$ can take arbitrary large values with non-zero probability) than $\sup(F)$ will be infinite.
-  \item $R(F)=\mu(F)+k\sigma (F)$, where $\sigma(F)$ is the standard deviation of $F$ and $k>0$ is a constant.
+  \item $R(F)=\mu(F)+k\sigma(F)$, where $\sigma(F)$ is the standard deviation of $F$ and $k>0$ is a constant.
 This risk measure is sometimes referred to as the safety margin.
-Optimisation using this risk measure is referred to as robust design optimisation22,23,24 and reflects the paradigms of a ‘$3\sigmaσ$-design’, a ‘$6\sigmaσ$-design’, etc.
+Optimisation using this risk measure is referred to as robust design optimisation22,23,24 and reflects the paradigms of a ‘$3\sigma$-design’, a ‘$6\sigma$-design’, etc.
   \item $R(F)=q_F(\alpha)$, where $q_F(\alpha)$ is the value of the quantile function, or simply the quantile, corresponding to a given probability $\alpha\in (0,1)$.
@@
 violating the constraint or underperforming, depending on the context) below a given threshold.
-The value $1−\alpha$ is the desired probability of failure.
+The value $1-\alpha$ is the desired probability of failure.
 Although rather intuitive, this choice has known computational issues which are discussed by Rockafellar and Royset27.
@@
 \begin{displaymath}
-  \mu (F)=E[F]=
+  \begin{array}{c}
+    \mu (F)=E[F]=\int F(\pmb x + \pmb u, \pmb y)p df(\pmb u, \pmb y)d\pmb u d\pmb y \\
+    \sigma^2(F)=E[(F-E[F])^2]=\\\int(F(\pmb x + \pmb u, \pmb y)-E[F])^2 p df(\pmb u, \pmb y)d\pmb u d\pmb y
+  \end{array}
 \end{displaymath}
@@
 \section{Numerical Example}
+\label{sec:num_example}
 
@@
 The engineering object to be optimised is shown in Fig.
-2 taken from Ref.
+\ref{fig:beam} taken from Ref.
 34.
 
-%picture
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.45\textwidth]{images/beam.png}
+    \caption{The cantilevered beam}
+    \label{fig:beam}
+\end{figure}
 
@@
 There are $N$ geometric constraints (the aspect ratios of each block (i.e.
-heights divided by widths) should not exceed 20) and $N$ constraints on the stress, calculated at the left end of each segment (stresses should not exceed  $\sigma=14000 N/cm^2$).
+heights divided by widths) should not exceed 20) and $N$ constraints on the stress, calculated at the left end of each segment (stresses should not exceed  $\bar{\sigma}=14000\; N/cm^2$).
 There is also also a constraint on the displacement at the tip, which should not exceed 2.5 cm.
-The load is $P = 50 000 N$, the Young's modulus is $E=2\cdot 10^7  N/cm^2$.
+The load is $P = 50 000\; N$, the Young's modulus is $E=2\cdot 10^7  N/cm^2$.
 
@@
 \begin{displaymath}
-  \mu (F)=E[F]=
+  \begin{array}{c}
+    y_0=y_0'=0, \\
+    y_i'=\frac{P\cdot l_i}{E\cdot I_i}\Big[ L+\frac{l_i}{2}-\sum\limits_{j=1}^i l_j\Big]+y_{i-1}', \\
+    y_i=\frac{P\cdot l_i^2}{2E\cdot I_i}\Big[L-\sum\limits_{j=1}^i l_j + \frac{2l_i}{3}\Big]+y_{i-1}'l_i+y_{i-1}.
+  \end{array}
 \end{displaymath}
@@
 \begin{displaymath}
-  \mu (F)=E[F]=
+  \sigma_i=\frac{M_i h_i}{2I_i}
 \end{displaymath}
@@
 \begin{displaymath}
-  \mu (F)=E[F]=
+  \begin{array}{c}
+    \min\limits_{\pmb b, \pmb h}V(\pmb b, \pmb h) \\
+    s.t.\;1.0\le b_i \le 10.0, \\
+    5.0 \le h_i \le 100.0, \\
+    y_N\le 2/5, \\
+    \sigma_i \le \bar{\sigma}=14000, \\
+    \frac{h_i}{b_i}\le 20.
+  \end{array}
 \end{displaymath}
 
-With $N=50$ segments (corresponding to 100 design variables) the SQP solution of this problem is $V = 63704.598 cm^3$.
-MAM obtained the solution $V = 63935.360 cm^3$, using 1500 function evaluations (as compared to almost 8000 evaluations used by SQP).
+With $N=50$ segments (corresponding to 100 design variables) the SQP solution of this problem is $V = 63704.598\; cm^3$.
+MAM obtained the solution $V = 63935.360\; cm^3$, using 1500 function evaluations (as compared to almost 8000 evaluations used by SQP).
 The number of points in a trust region used to build approximations was 150.
@@
 Now let us add uncertainty to the problem, adding Gaussian noise with zero mean and variance 0.1 to each design variable.
-Using the μ(F) + kσ(F) risk measure, we obtain the following problem statement:
+Using the $\mu(F) + k\sigma(F)$ risk measure, we obtain the following problem statement:
 \begin{displaymath}
-  \mu (F)=E[F]=
+  \begin{array}{c}
+    \min\limits_{\pmb b, \pmb h}\mu(V(\pmb b, \pmb h) + k\sigma V(\pmb b, \pmb h)) \\
+    s.t.\;1.0\le b_i \le 10.0, \\
+    5.0 \le h_i \le 100.0, \\
+    \mu(y_N) + k\sigma(y_N)\le 2.5, \\
+    \mu(\sigma_i)+k\sigma(\sigma_i)\le \bar{\sigma}=14000, \\
+    \mu(\frac{h_i}{b_i})+k\sigma(\frac{h_i}{b_i})\le 20.
+  \end{array}
 \end{displaymath}
@@
 Hopefully, this explanation eliminates the confusion.
-Optimising $\mu(F) + k\sigma(F)$ with $k = 3.0$ using SQP on original functions, we obtain the value of the deterministic objective (volume of the structure) at the robust optimum $V = 67598.049 cm^3$.
+Optimising $\mu(F) + k\sigma(F)$ with $k = 3.0$ using SQP on original functions, we obtain the value of the deterministic objective (volume of the structure) at the robust optimum $V = 67598.049\; cm^3$.
 To compute $\mu(F)$ and $\sigma(F)$ at each iteration, 1000 points were used.
@@
 
-Solving the same problem using the new version of MAM suitable for problems with uncertainty, we obtain the value of the deterministic objective (volume of the structure) at the robust optimum $V = 67799.335 cm^3$.
+Solving the same problem using the new version of MAM suitable for problems with uncertainty, we obtain the value of the deterministic objective (volume of the structure) at the robust optimum $V = 67799.335\; cm^3$.
 To obtain this result, MAM again used 150 points per trust region to build metamodels, which resulted in a total of 1500 function evaluations, which is orders of magnitude smaller than the number of points required if no metamodels are used.
@@
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2017-04-12 2017/EUROGEN/eurogen.tex
EUROGEN: finish refs section

@@
 
+The Moving Least Squares Method (MLSM) was proposed in Ref.
+\cite{LancasterSalkauskas1981} for smoothing and interpolation of scattered data and later used in the mesh-free form of the FEM\cite{Liszka1984}.
+As suggested in Ref.\cite{ChoiYounYang2001}, it can be used as a technique for metamodelling and used in MDO frameworks.
+The MLSM is a weighted least squares method where the weights depend on the Euclidian distance from a sample point to where the surrogate model is to be evaluated.
+The weight value for a certain sample point decays as the distance increases.
+Describing the weight decay with a Gaussian function tends to be the most useful option even though many others have been evaluated in Ref.
+\cite{ToropovSchrammSahaiJones2005}.
+As demonstrated in Ref.
+\cite{PolynkinToropov2010}, the cross-validated MLSM can be used both for design variable screening and for surrogate modelling.
+In order to create an efficient MDO framework for problems with disparate discipline attributes Ref.
+\cite{OllarToropovJones2014} extended the optimization approach of MAM to the use of local DOEs and MLS approximations built in different subspaces of the total design variable space corresponding to the individual disciplines.
+The subspaces are finally combined into the total design variable space in which the resulting MDO problem is solved.
+
 This paper presents a new development in the Multipoint Approximation Method that makes it capable of handling problems with uncertainty in design variables as well as in additional ‘environmental’ variables.
@@
 Responses are evaluated at the DoE points and approximations are built using the obtained values.
-Currently, the pool of approximation methods available in MAM consists of a metamodel assemblies15 and the moving least-squares metamodels16-19.
+Currently, the pool of approximation methods available in MAM consists of a metamodel assemblies\cite{PolynkinToropov2012} and the moving least-squares metamodels\cite{LancasterSalkauskas1981,Liszka1984,ChoiYounYang2001,ToropovSchrammSahaiJones2005}.
 Other metamodel types could be used as well.
@@
     The approximate problem (\ref{eq:problem_approx}) is solved using Sequential Quadratic Programming (SQP) and the solution of this problem   determines the centre of the next trust region.
-    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region10.
+    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region\cite{KeulenToropovMarkine1996}.
     \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $\pmb x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations).
@@
 As it has been already said, a risk measure is a mapping that transforms a random response $F(\pmb x+ \pmb u,\pmb y)$  into a deterministic function that can be optimized.
-Next we provide some examples of risk measures2,3.
+Next we provide some examples of risk measures\cite{RockafellarUryasev2000,SchillingsSchulz2015}.
 
@@
 This risk measure is sometimes referred to as the safety margin.
-Optimisation using this risk measure is referred to as robust design optimisation22,23,24 and reflects the paradigms of a ‘$3\sigma$-design’, a ‘$6\sigma$-design’, etc.
+Optimisation using this risk measure is referred to as robust design optimisation\cite{KochYangGu2004,Yao2011,ZhangZhuChenArendt2013} and reflects the paradigms of a ‘$3\sigma$-design’, a ‘$6\sigma$-design’, etc.
   \item $R(F)=q_F(\alpha)$, where $q_F(\alpha)$ is the value of the quantile function, or simply the quantile, corresponding to a given probability $\alpha\in (0,1)$.
-This choice corresponds to the paradigm of reliability-based design optimization25,23,26.
+This choice corresponds to the paradigm of reliability-based design optimization\cite{AouesChateauneuf2010,Yao2011,YounChoi2004}.
 In this approach, one wants to keep the probability of the system failing (i.e.
@@
 The value $1-\alpha$ is the desired probability of failure.
-Although rather intuitive, this choice has known computational issues which are discussed by Rockafellar and Royset27.
+Although rather intuitive, this choice has known computational issues which are discussed by Rockafellar and Royset\cite{RockafellarRoyset2010}.
 \end{itemize}
@@
 To compute them efficiently, methods of high-dimensional integration must be employed.
-Here, quasi-Monte-Carlo integration30 based on Niederreiter sequences31 is used.
+Here, quasi-Monte-Carlo integration\cite{Caflisch1998} based on Niederreiter sequences\cite{Niederreiter1992} is used.
 
@@
 
-The example considered in this study is a classical engineering optimisation problem known as the scalable cantilevered beam34.
+The example considered in this study is a classical engineering optimisation problem known as the scalable cantilevered beam\cite{Vanderpllaats2001}.
 The engineering object to be optimised is shown in Fig.
 \ref{fig:beam} taken from Ref.
-34.
+\cite{Vanderpllaats2001}.
 

### 2017-05-23 2017/NACO/paper.tex
Add a new version of Globalizer paper

% !!!IMPORTANT NOTE: Please read carefully all information including those preceded by % sign
%Before you compile the tex file please download the class file AIMS.cls from the following URL link to the
%local folder where your tex file resides. http://aimsciences.org/journals/tex-sample/AIMS.cls.
\documentclass{aims}
\usepackage{amsmath}
  \usepackage{paralist}
  \usepackage{graphics} %% add this and next lines if pictures should be in esp format
  \usepackage{epsfig} %For pictures: screened artwork should be set up with an 85 or 100 line screen
\usepackage{graphicx}  \usepackage{epstopdf}%This is to transfer .eps figure to .pdf figure; please compile your paper using PDFLeTex or PDFTeXify.
 \usepackage[colorlinks=true]{hyperref}
   % Warning: when you first run your tex file, some errors might occur,
   % please just press enter key to end the compilation process, then it will be fine if you run your tex file again.
   % Note that it is highly recommended by AIMS to use this package.
\hypersetup{urlcolor=blue, citecolor=red}
%\usepackage{hyperref}

\usepackage[utf8]{inputenc}

  \textheight=8.2 true in
   \textwidth=5.0 true in
    \topmargin 30pt
     \setcounter{page}{1}

% The next 5 line will be entered by an editorial staff.
\def\currentvolume{X}
 \def\currentissue{X}
  \def\currentyear{200X}
   \def\currentmonth{XX}
    \def\ppages{X--XX}
     \def\DOI{10.3934/xx.xx.xx.xx}

 % Please minimize the usage of "newtheorem", "newcommand", and use
 % equation numbers only situation when they provide essential convenience
 % Try to avoid defining your own macros

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}
\newtheorem*{main}{Main Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{conjecture}{Conjecture}
\newtheorem*{problem}{Problem}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}{Remark}
\newtheorem*{notation}{Notation}
\newcommand{\ep}{\varepsilon}
\newcommand{\eps}[1]{{#1}_{\varepsilon}}

%% Place the running title of the paper with 40 letters or less in []
 %% and the full title of the paper in { }.
\title[Globalizer: a supercomputer software] %Use the shortened version of the full title
      {Globalizer: a novel supercomputer software system for solving time-consuming global optimization problems}

% Place all authors' names in [ ] shown as running head, Leave { } empty
% Please use `and' to connect the last two names if applicable
% Use FirstNameInitial.  MiddleNameInitial. LastName, or only last names of authors if there are too many authors
\author[V. P. Gergel, K. A. Barkalov and A. V. Sysoyev]{}

% It is required to enter 2010 MSC.
\subjclass{Primary: 90C26; Secondary: 65Y05.}
% Please provide minimum  5 keywords.
 \keywords{Global optimization, information-statistical theory, parallel computing, high-performance
computer systems,  supercomputer technologies.}

% Email address of each of all authors is required.
% You may list email addresses of all other authors, separately.
 \email{victor.gergel@itmm.unn.ru}
 \email{konstantin.barkalov@itmm.unn.ru}
 \email{alexander.sysoyev@itmm.unn.ru}

% Put your short thanks below. For long thanks/acknowlegements,
%please go to the last acknowlegments section.
\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}

% Add corresponding author at the footnote of the first page if it is necessary.
% Plase add $^*$ adjacent to the corresponding author's name on the first page.
% The example shown in this template is if the first author is the corresponding author.
\thanks{$^*$ Corresponding author: Konstantin  Barkalov}

\begin{document}
\maketitle

% Enter the first author's name and address:

\centerline{\scshape Victor Gergel, Konstantin  Barkalov$^*$ and Alexander Sysoyev}
\medskip
{\footnotesize
 % please put the address of the second  and third author
 \centerline{Lobachevsky State University of Nizhny Novgorod}
   \centerline{23, Gagarin Avenue}
   \centerline{Nizhny Novgorod, Russia}
}

\bigskip

% The name of the associate editor will be entered by an editorial staff
% "Communicated by the associate editor name" is not needed for special issue.
 \centerline{(Communicated by the associate editor name)}

%The abstract of your paper
\begin{abstract}
In this paper, we describe the Globalizer software system for solving the global
optimization problems. The system is designed to maximize the use of computational
potential of the modern high-performance computational systems in order to solve the
most time-consuming optimization problems. The highly parallel computations are facilitated
using various distinctive computational schemes: processing several optimization
iterations simultaneously, reducing multidimensional optimization problems using multiple
Peano space-filling curves, and multi-stage computing based on the nested block reduction schemes.
These novelties provide for the use of the supercomputer system capabilities with shared
and distributed memory and with large numbers of processors to solve the global
optimization problems efficiently.
\end{abstract}

%The title of your section 1
\section{Introduction}\label{sec:intro}

Global (or multiextremal) optimization problems are among the most complex problems
in both theory and practice of optimal decision making. In these kinds of problems,
the criterion to be optimized has several local optima within the search domain,
which have different values. The existence of several local optima essentially makes
finding the global optimum difficult essentially, since it requires examining the whole feasible
search domain. The volume of computations for solving global optimization problems
can increase exponentially with increasing number of varied parameters.
\par
These global optimization problem features impose special requirements on the quality
of the optimization methods and on the software to implement these ones. The global
optimization methods should be highly efficient, and the software systems should be developed on
a good professional basis. In general, the global optimization problems can be solved at
a reasonable time by employing parallel computations on modern supercomputer systems only.
\par
The general state of the art in the field of global optimization has been presented in a
number of key monographs  \cite{floudasPardGO}, \cite{horstTuyGO}, \cite{locatelliSchoenGO}, \cite{pinterGO}, \cite{strSergGO}, \cite{zilinskTornGO}, \cite{zhigljavskyRandGO}, etc.
The development of optimization methods, which use the high-performance computational
systems to solve the time-consuming global optimization problems, is an area of
intensive research --- see, for instance, \cite{censorZeniosParGO}, \cite{ciegisHentyParGO},
\cite{luqueAlbaGA}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}. The obtained
theoretical results provide the efficient solutions of many applied global
optimization problems in various fields of scientific and technical applications \cite{Famularo1999}, \cite{fasanoPinter2013},
\cite{floudasPardalosGOState}, \cite{Kvasov2015}, \cite{Menniti}, \cite{locatelliSchoenGO},
\cite{luqueAlbaGA}, \cite{pardalosZhigljavskyZilinskas2016}, \cite{pinterGO}.
\par
At the same time, the practical implementation of these global optimization algorithms
within the framework of industrial software systems is quite limited. In many cases,
software implementations are experimental in nature and are used by the developers
themselves to obtain the results from the computational experiments required for the
scientific publications. This situation originates from high development costs of the
professional software systems, which can be used by numerous users. In addition, the global
optimization problems could be solved in an automatic mode rarely because of the
complexity of these ones. The user should actively control the global search
process that implies an adequate level of qualification in the field of optimization
(particularly, the user should know and understand the global optimization methods well).
\par
In the market of the global optimization software, one can select from the following systems:
\begin{itemize}
\item LGO (Lipschitz Global Optimization) \cite{pinterGO} was designed to solve the global
optimization problems, where the criteria and constraints satisfy the Lipschitz condition.
The system is a commercial product based on the diagonal extensions of the one-dimensional multiextremal optimization algorithms.
\item GlobSol \cite{kearfott2009} is oriented on solving the global optimization
problems as well as of the systems of the nonlinear equations. The system includes
the interval methods based on the branch and bound approach. There are some extensions
of the system for the parallel computations, and it is available for free usage.
\item LINDO \cite{linSchrage2009} is featured by a wide spectrum of optimization methods for solving linear, integer, stochastic, nonlinear, and global optimization problems. The ability to interact with the Microsoft Excel software environment is a useful feature of this system. The system is widely used in the practical applications and is available for free usage.
\item IOSO (Indirect Optimization on the basis of Self-Organization) \cite{iosoDescription} is oriented on solving a wide class of the optimization problems including the global optimization ones. The system is widely used to solve the applied problems in various application fields. There is a version of the system for exciting on the parallel computational systems. The system is a commercial product, but it is available for the trial usage.
\item MATLAB Global Optimization Toolkit \cite{venkataraman2009} contains a wide
spectrum of methods for solving the global optimization problems, including multistart
algorithms, global pattern search, simulated annealing methods, etc. The library is
compatible with the TOMLAB system \cite{holmstromEdvall2004}, which is an additional
MATLAB extension. It should also be noted that similar libraries for solving the global
optimization problems are available for MathCAD, Mathematica, and Maple systems as well.
\item BARON (Branch-And-Reduce Optimization Navigator) \cite{sahinidis1996} was designed to
solve continuous integer programming and global optimization problems using the branch and
bound approach. BARON is included in the GAMS (General Algebraic Modeling System) system \cite{bussieckMeeraus2004}.
\item Global Optimization Library in R is a large collection of optimization methods
implemented in the R language \cite{mullen2014}. Among these methods, stochastic and
deterministic global optimization algorithms and the branch and bound method were implemeted.
\end{itemize}
\par
The list provided above is certainly not exhaustive – additional information on the
software systems for solving a wider spectrum of optimization problems can be found, for example, in
\cite{mongeauKarsentyRouze2000}, \cite{pinter2009}, \cite{riosSahinidis2013}.
\par
In this paper, a novel Globalizer software system is presented. Globalizer expands the collection of the parallel optimization systems mentioned above. The Globalizer system was designed to solve the time-consuming multiextremal optimization problems that is an advantage of the system. In order to obtain the global optimized solutions within a reasonable time, the system exploits the huge computational potential of high-performance massively parallel systems.

The further structure of the paper is as follows. In Section \ref{sec:problem},
a statement of the multidimensional global optimization problem is given and an
approach for reducing this one to a one-dimensional optimization problem is described.
In Section \ref{sec:parallel}, the parallel computation schemes are discussed and
the parallel optimization methods are presented. In Section \ref{sec:arch}, the architecture of
the Globalizer system is described. In Section \ref{sec:experiments}, the results of the numerical experiments,
which confirm the parallel efficiency of the Globalizer system, are presented. Finally,
Section \ref{sec:concl} presents the conclusions and some ideas for the future research.

\section{Multidimensional Global Optimization Problems and Dimension Reduction}
\label{sec:problem}
In this paper, the core class\footnote{In general, the Globalizer can be applied for solving
multicriterial multiextremal multidimensional optimization problems with nonlinear constraints.}
of optimization problems, which can be solved using
Globalizer, is formulated. This class involves the multidimensional global
optimization problems without constraints, which can be defined in the following way:
\begin{equation}
\label{eq:task}
\begin{array}{cr}\\
  \varphi(y^*)=\min\{\varphi(y):y\in D\} \\
  D=\{y\in \mathbf{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
\end{array}
\end{equation}
It is supposed, that the objective function \(\varphi(y)\) satisfies the Lipschitz condition
\begin{equation}
\label{eq:lip}
|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
\end{equation}
where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbf{R}^N\) space.
\par
Usually, the minimized function \(\varphi(y)\) is defined as a computational procedure,
according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
(let us further call such a calculation \textit{a trial}). It is supposed that this procedure
is a time-consuming one. As a result, the overall time of solving the optimization
problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
is highly important, since an estimate of the global minimum can be constructed on the
base a finite number of computed values of the optimized function only in this case .
\par
As it has been shown earlier by many researchers
(see, for instance, [12], [25], [38], [51]),
finding the numerical estimate of the global optimum implies constructing a coverage of
the search domain \(D\). As a result, the computational costs of solving the global
optimization problems are readily very high even for a small number of varied parameters
(the dimensionality of the problem). A notable reduction in the volume of computations
can be achieved when the coverage of the search domain is non-uniform, i. e. the series
of trial points is only dense in a vicinity of the global optimum point. The construction
of such a non-uniform coverage could be provided in an adaptive way, when the selection
of the next trial points is determined by using the search information ((the preceding
trial points and the values of the minimized function at these points) obtained in the course
of computations. This necessary condition complicates considerably the computational schemes
for global optimization methods since it implies a complex analysis of a large amount
of multidimensional search information. As a result, many optimization algorithms use
various approaches to the dimensional reduction [38], [46], [48], [50], [51].
\par
Within the framework of the information-statistical global optimization theory,
the Peano space-filling curves (or evolvents) \(y(x)\) mapping the interval \([0,1]\)
onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
reduction [46], [48], [50], [51].
\par
As a result of the reduction, the initial multidimensional global optimization
problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
\begin{equation}
\label{eq:oneDimTask}
\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}
\end{equation}
\par
It is important to note that this dimensionality reduction scheme transforms the minimized
Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
function \(\varphi(y(x))\), which satisfy the uniform H{\"o}lder condition i. e.
\begin{equation}
\label{eq:holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
\end{equation}
where the constant H is defined by the relation \(H=4L\sqrt{N}\), \(L\) is the Lipschitz
constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem (\ref{eq:task}).
\par
The algorithms for the numerical construction of the Peano curve approximations are
presented in \cite{strSergGO}. As an illustration, an approximation of the Peano curve
for the third density level is shown in Figure \ref{fig:peanoC}. Figure \ref{fig:peanoC}
demonstrates the movement order in a two-dimensional domain to construct the Peano
curve approximation; the precision of the Peano curve approximation is determined by the
density level used in the construction.
\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{pictures/peanoC.eps}
    \caption{A Peano curve approximation for the third level}
    \label{fig:peanoC}
\end{figure}

\par
The computational scheme obtained as a result of the dimensionality reduction consists of the following
(see Figure \ref{fig:peanoCUsage}):
\begin{itemize}
  \item The optimization algorithm performs the minimization of the reduced one-dimensional
  function \(\varphi(y(x))\) from (\ref{eq:oneDimTask}),
  \item After determining the next trial point \(x\), a multidimensional image \(y\) in the mapping \(y(x)\) is calculated,
  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the multidimensional point \(y\in D\),
  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
\end{itemize}

\begin{figure}
    \centering
    \includegraphics[width=0.55\textwidth]{pictures/peanoCUsage.eps}
    \caption{The computational scheme for obtaining the value of the reduced one-dimensional function \(\varphi(y(x))\)}
    \label{fig:peanoCUsage}
\end{figure}

\section{Parallel Computations for Solving Global Optimization Problems}
\label{sec:parallel}
Let us consider the parallelization methods used widely in the theoretical and
practical application of parallel computing within the context of global optimization problems:
\begin{itemize}
  \item The distribution of the search domain \(D\) among the available computing units
  (data parallelization scheme). In the case of optimization problems, this approach is
  insufficient, since in organizing such computations the subdomain, which contain the
  sought global minimum, will be processed by only one processor, and, therefore,
  all of the remaining processors would perform the excess computations.
  \item The parallelization of the computational schemes for optimization algorithms
  (task parallelization scheme). This approach is also insufficient since the direct
  computational costs for executing optimization algorithms are relatively low
  (the majority of computations in the course of a global search are represented by
  calculations of the optimized function values due to the initial assumption of
  considerable computational costs for such calculations).
  \item The parallelization of the computations executed in order to obtain values
  for the optimized function. This approach can be applied since the most computationally
  costly part of the global search process will be parallelized. However, this method is not
  featured by the generality (the development of parallelization methods is to be performed
  every time from scratch while solving each particular optimization problem).
\end{itemize}
\par
Within the framework of information-statistical theory, a general approach to
parallelization computations when solving global optimization problems has been
proposed --- \textit{the parallelism of computations is provided by means of simultaneously
computing the values of the minimized function \(\varphi(y)\) at several different
points within the search domain \(D\)} – see, for example, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
This approach provides parallelizaion for the most costly part of computations in the global search process.
\par
The global optimization algorithms implemented in Globalizer will be described step-by-step below.
In Subsection \ref{subsec:corepar}, the core sequential multidimensional algorithm of global search (MAGS) will be presented.
In Subsection \ref{subsec:sharedpar}, a parallel generalization of the MAGS algorithm for parallel computations on
computer systems with shared memory will be described. In Subsection \ref{subsec:distribpar}, the scheme
for parallel computations by multiprocessor systems with distributed memory will be provided.

\subsection{Core Multidimensional Generalized Algorithm of Global Search}
\label{subsec:corepar}
The information-statistical theory of global search formulated in \cite{strongin1978}, \cite{strSergGO}
has served as a basis for the development of a large number of efficient multiextremal optimization methods --- see, for example, \cite{barkalovGergel2014}, \cite{Pizzuti}, \cite{gergel1996}, \cite{gergel1997}, \cite{grishaginStrongin1984}, \cite{sergeyev1995}, \cite{sergeyev1999}, \cite{sergeyevGrishagin1994}, \cite{sergeyevGrishagin2001}, \cite{sergeyevStronginLera2013}, \cite{Famularo2001} etc.
\par
Multidimensional Algorithm of Global Search (MAGS) established the basis for the
methods applied in Globalizer. The general computational scheme of MAGS can be
presented as follows --- see \cite{strongin1978}, \cite{strSergGO}.
\par
Let us introduce a simpler notation for the problem being solved on a computing node
\begin{equation}
\label{eq:oneDimFunc}
f(x) = \varphi(y(x)):x\in [0,1].
\end{equation}
\par
The initial iteration of the algorithm is performed at an arbitrary point \(x^1(0,1)\).
Let us assume further \(k\), \(k>1\), iterations of a global search are to be completed.
The selection of the trial point \(k+1\) for the next iteration is performed according to the following rules.
\par
\textit{Rule 1}. To renumber the points for the preceding trials by the lower indices in order of increasing value of coordinates
\begin{equation}
  \label{step1}
0=x_0<x_1<...<x_{k+1}=1.
\end{equation}
The points \(x_0\), \(x_k\) were introduced additionally for convenience in further explanation,
the values of the minimized function \(z_0\), \(z_k\) at these points are undefined.
\par
\textit{Rule 2}. To compute a current estimate of the H{\"o}lder constant \(H\) from (\ref{eq:holder})
\begin{equation} \label{step2}
m=\max_{1\leq i\leq k}\frac{|z_i-z_{i-1}|}{\rho_i}, \;
M = \left\{
   \begin{array}{lr}
     rm, & m > 0,\\
     1, & m = 0,
   \end{array}
  \right.
\end{equation}
as the maximum of the relative differences of the minimized function \(f(x)\) from (\ref{eq:oneDimFunc})
on the set of executed trial points \(x_i,1\leq i\leq k\) from (\ref{step1}). Further, \(\rho_i=(x_i-x_{i-1})^\frac{1}{N},1\leq i\leq k+1\).
The constant \(r\), \(r>1\), is the \textit{parameter} for the algorithm.
\par
\textit{Rule 3}. For each interval \((x_{i-1},x_i),1\leq i\leq k+1\), compute the characteristics \(R(i)\) where
\[
R(i)=2\rho_i-4\frac{z_i}{M},\quad i=1,
\]
\begin{equation} \label{step3}
R(i)=\rho_i+\frac{(z_i-z_{i-1})^2}{M^2\rho_i}-2\frac{z_i+z_{i-1}}{M},\quad 1<i<k+1, \\
\end{equation}
\[
R(i)=2\rho_{i}-4\frac{z_{i-1}}{M},\quad i=k+1.
\]

\par
\textit{Rule 4}. To determine the interval with the maximum characteristic
\begin{equation} \label{step4}
R(t)=\max_{1\leq i \leq k+1}R(i).
\end{equation}
\par
\textit{Rule 5}. To execute a new trial (computation of the minimized function value \(f(x)\))
at the point \(x^{k+1}\) located within the interval with the maximum characteristic from (\ref{step4})
\begin{equation} \label{step5}
  x_{k+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{|z_{t}-z_{t-1}|}{M}\right]^N,\; \textrm{ if } 1<t<k+1,
\end{equation}
\[
  x_{k+1}=\frac{x_t+x_{t-1}}{2},\; \textrm{ if } t=1 \textrm{ or } t=k+1.
\]

\par
The stop condition by which the trials are terminated, is defined by the inequality
\begin{equation}
  \label{eq:stop_1}
\rho_t<\varepsilon
\end{equation}
for the interval with the maximum characteristic from (\ref{step4}) and \(\varepsilon >0\) is the predefined
accuracy of the problem solution. If the stop condition is not satisfied, the index \(k\) is incremented by unity,
and a new global search iteration is executed.
\par
In order to explain the algorithm presented above, let us note the following.
The characteristics \(R(i), 1\leq i\leq k+1\), calculated according to (\ref{step3})
could be interpreted as some measures of importance to the intervals with respect to
location of the global minimum point. Thus, the scheme for selecting the interval
for executing the next trial described in (\ref{step4}), (\ref{step5}) becomes more clear --- the point of every next
global search iteration is selected within the interval where the global minimum point can most likely be found.
\par
The conditions for the algorithm’s convergence described above were examined, for example, in \cite{strSergGO}.

\subsection{Parallel Computations for Systems with Shared Memory}
\label{subsec:sharedpar}
Modern supercomputer computational systems consist of plenty of computer nodes,
which include several multicore processors. At that, random access memory is
shared for the computing nodes --- the content of any memory element can be read (written)
by any computer cores available at any arbitrary moment. In most cases, shared memory is
uniform --- the time characteristics for accessing memory are the same for all computer
cores and for all memory elements.
\par
The following speculations can form the basis for selecting parallel computation
organization methods. As was mentioned above when describing the MAGS algorithm,
the characteristics \(R(i),1\leq i\leq k+1\), calculated according to (\ref{step3})
can be interpreted as some measures of the importance of intervals with respect to the location of
the global minimum point. Following this understanding, each successive global search
iteration is executed within the most important interval with the maximum value of the
characteristic. So far, in this case it is obvious how to select the other intervals for
organizing simultaneous computations of the minimized function values at several different
points within the search domain as well --- these should be the intervals with the next
magnitudes of characteristics.
\par
The computational scheme of the Parallel Multidimensional Algorithm of Global Search for
computer systems with shared memory (PMAGS) was developed based the approach considered
above and is practically identical to the MAGS scheme --- the differences consist just in the
following set of rules --- see \cite{strSergGO}, \cite{stronginGergelBarkalovParGO}.
\par
\textit{Rule 4'}. To arrange the characteristics of the intervals obtained according to (\ref{step3}) in decreasing order
\begin{equation}
\label{step4par}
R(t_1)\geq R(t_2)\geq \dots \geq R(t_{k})\geq R(t_{k+1})
\end{equation}
and to select \(p\) intervals with the indices \(t_j,1\leq j\leq p\), with
the highest values of characteristics (\(p\) is the number of processors/cores used for the parallel computations).
\par
\textit{Rule 5'}. To execute new trials (computations of values for the minimized
function \(f(x)\)) at the points \(x_{k+j},1\leq j\leq p\), located in the
intervals with the highest characteristics from (\ref{step4par})

\begin{equation} \label{step5par}
x_{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\; \textrm{ if } 1<t_j<k+1,
\end{equation}
\[
x_{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\; \textrm{ if } t_j=1 \textrm{ or } t_j=k+1.
\]
\par
The stop condition for the algorithm (\ref{eq:stop}) should be checked for all
intervals from (\ref{step4par}), in which the scheduled trials are executed
\begin{equation}
  \label{eq:stop}
\rho_{t_j}<\varepsilon,1\leq j\leq p.
\end{equation}
As before, if the stop condition is not fulfilled, the index \(k\) is
incremented by \(p\), and a new global search iteration is executed.
\par
The conditions for the developed parallel algorithm’s convergence and for the
non-redundancy of parallel computations have been examined in \cite{strSergGO}.
Thus, in particular, when the conditions for convergence are satisfied, the parallel
computations are non-redundant as compared to the sequential method when using up to
\(2^N\) processors/cores (where \(N\) is the dimensionality of the global optimization problem being solved).

\subsection{Parallel Computations for Systems with Distributed Memory}
\label{subsec:distribpar}
The next level of parallel computations in high-performance computer systems consists of
using several computational nodes. For that, each computational node has its own separate
memory and, in this case, the interaction between different nodes can be provided by
means of data transfer only via the computer communication network.
\par
To organize parallel computations for multiprocessor systems with distributed memory, using a set of mappings
\begin{equation}
  \label{eq:maps}
Y_s(x)=\{y^1(x),\dots,y^s(x)\}
\end{equation}
instead of applying a single Peano curve \(y(x)\) has been proposed in \cite{strongin1992}, \cite{strSergGO}.
In order to build the set \(Y_s(x)\), several different approaches can be applied.
Thus, for example, in \cite{strongin1992} a scheme has been applied by which each
mapping \(y_i(x)\) from \(Y_s(x)\) is built as the result of some shift along the main
diagonal of the hypercube \(D\). This way, the set of constructed Peano curves enables the
close prototypes \(x'\), \(x''\)  to be obtained from the interval \([0,1]\) for any close
multidimensional images \(y'\), \(y''\) from \(D\) differing in one coordinate,
only for some mapping \(y_k(x), 1\leq k\leq s\).
\par
Some other methods for constructing multiple mappings have been examined in \cite{stronginGergelBarkalovParGO}.
\par
The set of mappings \(Y_s(x)\) from (\ref{eq:maps}), for multidimensional problem (\ref{eq:task})
generates \(s\) subsidiary information-linked one-dimensional problems of the kind (\ref{eq:oneDimTask}):
\begin{equation}
  \label{eq:oneDimProblemSerie}
  \varphi(y^l(x^*))=\min\{\varphi(y^l(x)):x\in [0,1]\},1\leq l\leq s.
\end{equation}
\par
It is important to note that the family of one-dimensional problems
\[
\varphi(y^l(x)),1 \leq l \leq s,
\]
obtained as a result of the dimensional reduction is information-linked --- the function
values computed for any problem \(\varphi(y^l(x))\) from the family (\ref{eq:oneDimProblemSerie}) can be used for all
of the remaining problems of this family.
\par
The information compatibility of the problems from the family (\ref{eq:oneDimProblemSerie}) allows the parallel
computations to be organized in an obvious enough way. Thus, each particular problem can
be solved by a separate processor in the computing system; the exchange of search information
between the processors should be performed during the course of the computations.
As a result of using such a parallelization method, a unified approach for organizing parallel
computations for a multiprocessor computer systems with distributed, as well as shared
memory can be developed. This method consists of the following.
\begin{enumerate}
  \item The family of reduced one-dimensional information-linked problems (\ref{eq:oneDimProblemSerie}) is
  distributed among the computational nodes of a multiprocessor system. One problem,
  as well as several others from within the family, can be allocated to each particular computational node.
  \item The Parallel Multidimensional Algorithm of Global Search (PMAGS, see Subsection \ref{subsec:sharedpar})
  is applied to solve the allocated problems from the family (\ref{eq:oneDimProblemSerie})
  at each computational node, supplemented by the following rules of information interaction.
  \begin{enumerate}
    \item Prior to beginning a new trial at any point \(x'\in [0,1]\) for any problem \(\varphi(y_l(x)),1\leq l\leq s\),
    the following should be performed:
    \begin{itemize}
      \item compute the image \(y'\in D\) for the point \(x'\in [0, 1]\) according to the mapping \(y_l(x)\),
      \item determine  the prototypes \(x_i',1\leq i\leq s\), for each
      problem from the family (\ref{eq:oneDimProblemSerie}),
      \item send the prototypes \(x_i',1\leq i\leq s\) to all computational
      nodes employed in order to exclude the repeated selection of the intervals in
      which the prototypes fall, for using them to determine the points for new trials.
      To organize the data transfer, a queue can be formed at each computational
      node for sending the trial points and receiving the minimized function values at these points.
    \end{itemize}
    \item After completing any trial for any problem \(\varphi(y_l(x)),1\leq l\leq s\), at any point
    \(x'\in[0,1]\), it is necessary:
    \begin{itemize}
      \item to determine all prototypes \(x_i',1\leq i\leq s\), for the point of the completed trial
      for each problem from the family (\ref{eq:oneDimProblemSerie}),
      \item to send the prototypes \(x_i',1\leq i\leq s\), and the result of the
      trial \(z'=\varphi(y_l(x'))\) to all computational nodes employed to include the
      obtained data into the search information processed within the rules of the parallel global search algorithm.
    \end{itemize}
    \item Prior to starting the next global search iteration, the algorithm should
    check the queue of the received data; if there are some data in the queue,
    they should be included in the search information.
\end{enumerate}
\end{enumerate}
\par
The possibility of asynchronous data transfer (the computational nodes process received
data only upon receipt) is a meaningful feature of such a scheme for organizing parallel computations.
\par
In addition, there is no single managing node within this scheme. The number of
computational nodes can vary during the course of the global search, and excluding any
node does not result in the loss of the sought global minimum of the minimized multiextremal function.
\par
This computational scheme generates the Generalized Parallel Multidimensional Algorithm of
Global Search (GPMAGS) for high-performance computing systems, which may include many
multiprocessor (multicore) computational nodes with distributed memory as well as
accelerators for computations based on Intel Xeon Phi processors and on general purpose graphic processors.
\par
Additional information on organizing such parallel computation schemes is provided in \cite{gergelSidorov2015}.

\section{Globalizer System Architecture}
\label{sec:arch}
As previously mentioned in Section \ref{sec:intro}, the development of software systems
for global optimization is complex professional work. The optimization systems should
include advanced optimization methods, provide for the accumulation and efficient
processing of search information, include various forms of visualization for the results,
provide tools for interacting with users, etc. All the features pointed out above
become much more complicated when organizing parallel computations using high-performance
multiprocessor systems of varying architectures.
\par
The Globalizer considered in this paper expands the family of global optimization
software systems successively developed by the authors during the past several years.
One of the first developments was the SYMOP multiextremal optimization system \cite{gergel1993},
which has been successfully applied for solving many optimization problems. A special
place is occupied by the ExaMin system (see, for example, \cite{barkalovGergel2015}),
which was developed and used extensively to investigate the application of novel parallel
algorithms to solve global optimization problems using high-performance multiprocessor computing systems.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{pictures/globalizerScheme.eps}
    \caption{The Globalizer system architecture (Blocks 1-2, 5-7 have been implemented; Blocks 3-4 and 8-11 are under development)}
    \label{fig:globalizerScheme}
\end{figure}

\par
The Globalizer system architecture is presented in Figure \ref{fig:globalizerScheme}. The structural components are:
\begin{itemize}
  \item Block 0 consists of the procedures for computing the function values (criteria
  and constraints) for the optimization problem being solved; this is an external
  block with respect to the Globalizer system and is provided by the user per a predefined interface.
  \item Blocks 1-4 form the optimization subsystem and solve the global optimization
  problems (Block 1), nonlinear programming (Block 2), multicriterial optimization (Block 3),
  and general decision making problems (Block 4). It is worth noting the sequential
  scheme of interaction between these components --- the decision making problems are
  solved using the multicriterial optimization block, which, in turn, uses the nonlinear programming block, etc.
  \item Block 5 is a subsystem for accumulating and processing the search information;
  this is one of the basic subsystems --- the amount of search information for complex
  optimization problems may appear to be quite large on the one hand, but, on the other
  hand, the efficiency of the global optimization methods depends to a great extent on how
  completely all of the available search data is utilized.
  \item Block 6 contains the dimensional reduction procedures based on Peano evolvents;
  the optimization blocks solve the reduced (one-dimensional) optimization problems;
  Block 6 provides interaction between the optimization blocks and the initial
  multidimensional optimization problem.
  \item Block 7 organizes the choice of parallel computation schemes in the Globalizer
  system subject to the computing system architecture employed (the numbers of cores in the
  processors, the availability of shared and distributed memory, the availability of
  accelerators for computations, etc.) and the global optimization methods applied.
  \item Block 8 is responsible for managing the parallel processes when performing the
  global search (determining the optimal configuration of parallel processes, distributing
  the processes between computer elements, balancing computational loads, etc.).
  \item Block 9 is a common managing subsystem, which fully controls the whole
  computational process when solving global optimization problems.
  \item Block 10 is responsible for organizing the dialog interaction with users for
  stating the optimization problem, adjusting system parameters (if necessary), and
  visualizing and presenting the global search results.
  \item Block 11 is a set of tools for visualizing and presenting the global search results;
  the availability of tools for visually presenting the computational results enables
  the user to provide efficient control over the global optimization process.
\end{itemize}
\par
The Globalizer system comes in two different options:
\begin{itemize}
  \item Globalizer Pro is a research version of the system providing a full set of
  capabilities that are oriented toward solving of the most complicated global optimization problems.
  \item Globalizer Lite is a limited version of the Globalizer system accessible for free for
  solving global optimization problems with a medium level of difficulty, and is accessible free of charge.
\end{itemize}

\section{Results of Numerical Experiments}
\label{sec:experiments}
In order to confirm the operability of the Globalizer system, results from two
series of computational experiments are presented in this section.
\par
The computational experiments were conducted using the Lobachevsky supercomputer at the
State University of Nizhny Novgorod (operation system --- CentOS 6.4, management system --- SLURM).
Each supercomputer node had 2 Intel Sandy Bridge E5-2660 processors, 2.2 GHz, with 64 GB RAM.
Each processor had 8 cores (i.e. a total of 16 CPU cores were available at each node).
In order to generate the executable program code, the Intel C++ 14.0.2 compiler was used.
\par
A number of nodes in the computer system utilized Intel Xeon Phi 5110P processors,
each containing 60 cores (a total of 240 threads). In addition, the supercomputer also
included computer nodes equipped with two GPU NVIDIA Kepler K20X, each providing 14
stream multiprocessors (a total of 2688 CUDA-cores). In order to generate the executable
code, the CUDA Toolkit 6.0 was used.
\par
The problems generated by the GKLS-generator \cite{gavianoKvasovLeraSergeev2003}
were selected for the test problems. This generator allowed for obtaining the multiextremal
optimization problems with \textit{a priori} known properties: the number of local minima,
the sizes of their attractors, the global minimum point, the function value at this point, etc.
In order to simulate the computation costs inherent in the applied optimization problems (see, for example, \cite{Barkalov2013}, \cite{gergel2013}, \cite{gergel2015}), in all of the experiments computing the objective function was complicated by auxiliary computations, not altering the form of the function and the position of its minima.
\par
The computational experiments considered below were first carried out using the ExaMin system (see \cite{barkalovGergel2015}, \cite{barkalovGergelLebedev2015}, \cite{barkalovGergelLebedevSysoev2015}, \cite{gergelLebedev2015}) and then were reproduced using the Globalizer system.

\subsection{Comparison with over methods}

The results of numerical comparison of three sequential algorithms -- DIRECT \cite{Jones}, DIRECT\textit{l} \cite{Gablonsky} and core global search algorithm (GSA) from section \ref{subsec:corepar} -- are provided below (results of the work of the first two algorithms are given in paper \cite{gavianoKvasovLeraSergeev2003}). Numerical comparison was carried out on function classes \textit{Simple} and \textit{Hard} of dimension 4 and 5 from \cite{gavianoKvasovLeraSergeev2003} since solving a problems of dimension 2 and 3 demands a small number of iterations. Global minimizer $y^\ast$ was considered to be found, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^\ast\right\|\leq\delta$. The size of the vicinity was selected (according to \cite{gavianoKvasovLeraSergeev2003}) as $\delta = \left\|b-a\right\|\sqrt[N]{\Delta}$, $N$ -- problem dimension, $a$ and $b$ -- borders of search domain $D$, parameter $\Delta=10^{-6}$ at $N=4$ and $\Delta=10^{-7}$ at $N=5$. When using the GSA for the class \textit{Simple} parameter $r=4.5$ was selected, for the class \textit{Hard} $r=5.6$ was selected; evolvent construction parameter was fixed as $m=10$. The maximum allowable number of iterations was $K_{max} = 10^6$.

The average number of iterations $k_{av}$ performed by the method for solving a series of problems from these classes is shown in table \ref{tab:3}. Symbol ''$>$''  reflects a situation in which not all problems of the class were solved by the method. This means that the algorithm was stopped as the maximum allowable number of iterations $K_{max}$ was achieved. In this case, value $K_{max}=10^6$ was used for calculating the average value of the number of iterations $k_{av}$ that corresponds to the lower estimate of this average value. The number of unsolved problems is specified in brackets.

\begin{table}
  \caption{Average number of iterations $k_{av}$}
  \label{tab:3}
  \center
  \begin{tabular}{lllll}
    \hline\noalign{\smallskip}
     $N$ & Problem class & DIRECT & DIRECT\textit{l} & GSA \\
    \noalign{\smallskip} \hline \noalign{\smallskip}
      4 &	\textit{Simple}	& $>$47282(4) &	18983 &	11953 \\
        & \textit{Hard} &	$>$95708(7) &	68754 &	25263 \\
      5	& \textit{Simple} &	$>$16057(1) &	16758 &	15920 \\
        & \textit{Hard} &	$>$217215(16) &	$>$269064(4) & $>$148342(4) \\
    \noalign{\smallskip}\hline
  \end{tabular}
\end{table}

Table \ref{tab:3} shows that GSA outperforms DIRECT and DIRECT\textit{l} methods on all classes of problems in terms of average number of iterations. In class 5-\textit{Hard} all three methods failed to solve all problems: DIRECT did not solve 16 problems, DIRECT\textit{l} and GSA, 4 problems.

\subsection{Parallel Computations Using Intel Xeon Phi processors}
The results of the numerical experiments with GPMAGS on an Intel Xeon Phi are provided in Table \ref{tab:exResults}.
The computations were performed using the \textit{Simple} and \textit{Hard} function classes with the
dimensions equal to 4 and 5. The parameters \(\Delta=10^{-6}\) at \(N=4\) and \(\Delta=10^{-6}\) at \(N=5\)
were used. When using the GPMAGS method, the parameter \(r=4.5\) was selected for the
\textit{Simple} class, and \(r=5.6\) was selected for the \textit{Hard} class.
\par
In the first series of experiments, serial computations using MAGS were executed.
The average number of iterations performed by the method for solving a series of
problems for each of these classes is shown in row I. The symbol ``\(>\)'' reflects the
situation where not all problems of a given class were solved by a given method.
It means that the algorithm was stopped once the maximum allowable number of
iterations \(K_{max}\) was achieved. In this case, the \(K_{max}\) value was used for
calculating the average number of iterations corresponding to the lower estimate of
this average value. The number of unsolved problems is specified in brackets.
\par
In the second series of experiments, parallel computations were executed on a CPU.
The relative ``speedup'' in iterations achieved is shown in row II; the speedup of
parallel computations was measured in relation to the sequential computations (\(p=1\)).

\begin{table}
  \centering
  \label{tab:exResults}
  \caption{Results of numerical experiments}
  \begin{tabular}{cccccccc}
    \cline{3-8}\noalign{\smallskip}
    \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}   \\
    \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
    \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
    \noalign{\smallskip}\hline
    I &
    \parbox{0.25\textwidth}{
    \begin{center}
    \textbf{Serial computations.}\\ \textit{Average number} \\ \textit{of iterations.}
    \end{center}		}
      & 1 & 11953 & 25263 & & 15920 & \(>\)148342 (4)  \\
    \hline \noalign{\smallskip}
II  & \textbf{Parallel computations}  %\multirow{3}{*}{}
  & 2 & 2.51 & 2.26 & & 1.19 & 1.36 \\
& \textbf{of CPU}. & 4 & 5.04 & 4.23 & & 3.06 & 2.86 \\
& \textit{Speedup} & 8 & 8.58 & 8.79 & & 4.22 & 6.56 \\
    \noalign{\smallskip}\hline	\noalign{\smallskip}
III & \textbf{Parallel computations} %\multirow{3}{*}{}
  & 60  & 8.13 & 7.32 & & 9.87 & 6.55  \\
&  o\textbf{f Xeon Phi}. & 120 & 16.33 & 15.82 & & 15.15 & 17.31 \\
& \textit{Speedup} & 240 & 33.07 & 27.79 & & 38.80 & 59.31 \\

    \noalign{\smallskip}\hline
  \end{tabular}
\end{table}

\par
The final series of experiments was executed using a Xeon Phi. The results of these
computations are shown in row III; in this case, the speedup factor is calculated in
relation to the PMAGS results on a CPU using eight cores (\(p=8\)). It can be noted that
relatively high speedup was achieved. For example, solving a problem from the 5-\textit{Hard} class
required on average only 633 parallel iterations on the Xeon Phi, whereas the number of
iterations was more than 37 thousand when using all of the CPU’s computing cores.

\subsection{Parallel Computations Using General Purpose Graphic Processors}
In the first series of experiments, only one graphic accelerator was used for solving
100 6-diminsional problems. The method’s reliability parameter \(r=4.5\) was selected.
\par
The average problem solving time using a GPU was 10.78 sec. whereas the averaged problem
solving time using all of the node’s 16 CPU cores was 53.8 sec.; an almost 5-fold improvement was observed.
\par
A larger scale computational experiment was carried out for the problems with
dimensions of \(N=8\) and \(N=10\): a total of 12 nodes within a cluster with 36
graphic accelerators (three accelerators per node) were employed, therefore, a total
of 96768 CUDA-cores were employed.
\par
The average time to solve the 8-dimensional problem was 405.6 seconds, 5.9 times
faster than the CPU version of the algorithm. The average time to solve the
10-dimensional problem was 2,055.8 sec. The CPU was not used to solve the 10-dimensional
problem because of the high degree of complexity in the computations.

\section{Conclusions} \label{sec:concl}

In this paper, the Globalizer global optimization system for solving the time-consuming global optimization problems using a wide spectrum of high-performance computational systems (the systems with shared and distributed memory, the systems with NVIDIA graphic processors and Intel Xeon Phi coprocessors) was considered. The highly parallel computations were facilitated using various distinctive computational schemes: processing several optimization iterations simultaneously, reducing multidimensional optimization problems using multiple Peano space-filling curves, and the multi-stage computing based on nested block reduction schemes. The numerical experiments have confirmed the efficiency of the Globalizer system.

In future, it is planned to extend the Globalizer system for solving the multicriterial multiextremal multidimensional optimization problems with the nonlinear constraints. As a valuable part of further work, Globalizer would be used for solving the applied optimization problems from various application fields.
% You may incorporate your references as follows in your main tex file.
% Using BibTex is not recommended but can be handled.
%\bibliographystyle{plain}
%\bibliography{paper}

\begin{thebibliography}{10}

\bibitem{barkalovGergel2014}
\newblock K. A. Barkalov and V. P. Gergel,
\newblock Multilevel scheme of dimensionality reduction for parallel global search algorithms,
\newblock in \emph{Proceedings of the 1st International Conference on
  Engineering and Applied Sciences Optimization}, (2014), 2111--2124.

\bibitem{barkalovGergel2015}[10.1007/s10898-016-0411-y]
\newblock K. Barkalov and V. Gergel,
\newblock Parallel global optimization on GPU,
\newblock \emph{J. Glob. Optim.}, \textbf{66(1)} (2016), 3--20.

\bibitem{barkalovGergelLebedev2015}[10.1007/978-3-319-21909-7_31]
\newblock K. Barkalov, V. Gergel and I. Lebedev,
\newblock Use of XeonPhi coprocessor for solving global optimization problems,
\newblock \emph{LNCS}, \textbf{9251} (2015), 307--318.

\bibitem{barkalovGergelLebedevSysoev2015}
\newblock K. Barkalov, V. Gergel, I. Lebedev, A. Sysoev, %http://ceur-ws.org/Vol-1482/
\newblock Solving the global optimization problems on heterogeneous cluster systems,
\newblock in \emph{CEUR Workshop Proceedings}, \textbf{1482} (2015), 411--419.
%\newblock (In Russian).

\bibitem{Barkalov2013}[10.1007/978-3-642-39958-9_14]
\newblock K. Barkalov, A. Polov inkin, I. Meyerov, S. Sidorov, N. Zolotykh,
\newblock SVM regression parameters optimization using parallel global search algorithm,
\newblock \emph{LNCS}, \textbf{7979} (2013), 154--166.

\bibitem{bussieckMeeraus2004}[10.1007/978-1-4613-0215-5_8]
\newblock M. R. Bussieck and A. Meeraus,
\newblock General algebraic modeling system (GAMS), in \emph{Modeling Languages in Mathematical Optimization} (ed. J. Kallrath),
\newblock Springer (2004), 137--157.

\bibitem{censorZeniosParGO}
\newblock Y. Censor and S. A. Zenios,
\newblock \emph{Parallel optimization: theory, algorithms, and applications},
\newblock Oxford University Press (1998).

\bibitem{ciegisHentyParGO}[10.1007/978-0-387-09707-7]
\newblock R. Ciegis, D. Henty, B. Kågström and J. Zilinskas,
\newblock \emph{Parallel scientific computing and optimization: advances and applications},
\newblock Springer (2009).

\bibitem{iosoDescription}
\newblock I. N. Egorov, G. V. Kretinin, I. A. Leshchenko and S. V. Kuptzov,
\newblock IOSO optimization toolkit --- novel software to create better design,
\newblock in \emph{9th AIAA/ISSMO Symposium on Multidisciplinary Analysis and Optimization}, 2002.
\newblock Available from \url{http://www.iosotech.com/text/2002\_4329.pdf}.

\bibitem{Famularo1999}[10.1016/S0005-1098(99)00058-8]
\newblock D. Famularo and P. Pugliese and Ya. D. Sergeyev,
\newblock A global optimization technique for checking parametric robustness,
\newblock \emph{Automatica}, \textbf{35} (1999), 1605--1611.

\bibitem{fasanoPinter2013}[10.1007/978-1-4614-4469-5]
\newblock G. Fasano and J. D. Pinter,
\newblock \emph{Modeling and optimization in space engineering},
\newblock Springer, 2013.

\bibitem{floudasPardalosGOState}[10.1007/978-1-4613-3437-8]
\newblock C. A. Floudas and M. P. Pardalos,
\newblock \emph{State of the art in global optimization: computational methods and applications},
\newblock Kluwer Academic Publishers, Dordrecht (1996).

\bibitem{floudasPardGO}[10.1007/s10898-008-9332-8]
\newblock C. A. Floudas and M. P. Pardalos,
\newblock \emph{Recent advances in global optimization},
\newblock Princeton University Press (2016).

\bibitem{Gablonsky}[10.1023/A:1017930332101]
\newblock J. M. Gablonsky and C. T. Kelley,
\newblock A locally-biased form of the DIRECT algorithm,
\newblock \emph{J. Glob. Optim.}, \textbf{21(1)} (2001), 27--37.

\bibitem{gavianoKvasovLeraSergeev2003}[10.1145/962437.962444]
\newblock M. Gaviano and D.E. Kvasov and D. Lera and Ya.D. Sergeev,
\newblock Software for generation of classes of test functions with known local and global minima for global optimization,
\newblock \emph{ACM Trans. Math. Software}, \textbf{29(4)} (2003), 469--480.

\bibitem{gergelLebedev2015}[10.1016/j.procs.2015.11.008]
\newblock V. Gergel and I. Lebedev,
\newblock Heterogeneous parallel computations for solving global optimization problems,
\newblock \emph{Procedia Comput. Sci.}, \textbf{66} (2015), 53--62.

\bibitem{gergel1993}
\newblock V. P. Gergel,
\newblock A software system for multi-extremal optimization,
\newblock \emph{Eur. J. Oper. Res.}, \textbf{65(3)} (1993), 305--313.

\bibitem{gergel1996}
\newblock V. P. Gergel,
\newblock A method for using derivatives in the minimization of multiextremum functions,
\newblock \emph{Comput. Math. Math. Phys.}, \textbf{36(6)} (1996), 729--742.

\bibitem{gergel1997}[10.1023/A:1008290629896]
\newblock V. P. Gergel,
\newblock A global optimization algorithm for multivariate functions with lipschitzian first derivatives,
\newblock \emph{J. Glob. Optim.}, \textbf{10(3)} (1997), 257--281.

\bibitem{gergel2013}[doi:10.1016/j.procs.2013.05.164]
\newblock V. P. Gergel, et al.,
\newblock High performance computing in biomedical applications,
\newblock \emph{Procedia Computer Science}, \textbf{18} (2013), 10--19.

\bibitem{gergel2015}[10.15866/ireaco.v8i1.4935]
\newblock V. P. Gergel, et al.,
\newblock Recognition of surface defects of cold-rolling sheets based on method of localities,
\newblock \emph{International Review of Automatic Control }, \textbf{8(1)} (2015), 51--55.

\bibitem{gergelSidorov2015}[10.1007/978-3-319-21909-7_49]
\newblock V. P. Gergel and S. V. Sidorov,
\newblock A two-level parallel global search algorithm for solving computationally intensive multi-extremal optimization problems,
\newblock \emph{LNCS}, \textbf{9251} (2015), 505--515.

\bibitem{grishaginStrongin1984}
\newblock V. A. Grishagin and R. G. Strongin,
\newblock Optimization of multi-extremal functions subject to monotonically unimodal constraints,
\newblock \emph{Engineering Cybernetics}, \textbf{5} (1984), 117--122.

\bibitem{holmstromEdvall2004}[10.1007/978-1-4613-0215-5_19]
\newblock K. Holmström and M. M. Edvall,
\newblock The TOMLAB optimization environment,
\newblock \emph{Modeling Languages in Mathematical Optimization},
\newblock 369--376, Springer (2004).

\bibitem{horstTuyGO}[10.1007/978-3-662-03199-5]
\newblock R. Horst and H. Tuy,
\newblock \emph{Global optimization: deterministic approaches},
\newblock Springer-Verlag, Berlin (1990).

\bibitem{Jones}[10.1007/BF00941892]
\newblock D. R. Jones, C. D. Perttunen and B. E. Stuckman,
\newblock Lipschitzian optimization without the Lipschitz constant,
\newblock \emph{J. Optim. Theory Appl.}, \textbf{79(1)} (1993), 157--181.

\bibitem{kearfott2009}[10.1080/10556780802614051]
\newblock R. B. Kearfott,
\newblock GlobSol user guide,
\newblock \emph{Optim. Methods Softw.}, \textbf{24} (2009), 687--708.

\bibitem{Kvasov2015}[10.1016/j.advengsoft.2014.09.014]
\newblock D. E. Kvasov and Ya. D. Sergeyev,
\newblock Deterministic approaches for solving practical black-box global optimization problems,
\newblock \emph{Adv. Eng. Softw.}, \textbf{80} (2015), 58--66.

\bibitem{Menniti}[10.1016/j.epsr.2007.10.009]
\newblock D. E. Kvasov, D. Menniti, A. Pinnarelli, Ya. D. Sergeyev and N. Sorrentino,
\newblock Tuning fuzzy power-system stabilizers in multi-machine systems by global optimization algorithms based on efficient domain partitions,
\newblock \emph{Electric Power Systems Research}, \textbf{78(7)} (2008), 1217--1229.

\bibitem{Pizzuti}[10.1007/s00211-002-0419-8]
\newblock D. E. Kvasov, C. Pizzuti and Ya. D. Sergeyev,
\newblock Local tuning and partition strategies for diagonal go methods,
\newblock \emph{Numerische Mathematik}, \textbf{94(1)} (2003), 93--106.

\bibitem{liberti2009}[10.1007/0-387-30528-9_8]
\newblock L. Liberti,
\newblock Writing global optimization software,
\newblock in \emph{Nonconvex Optimization and Its Applications }, \textbf{84}, 211--262,
\newblock Springer (2006).

\bibitem{linSchrage2009}[10.1080/10556780902753221]
\newblock Y. Lin and L. Schrage,
\newblock The global solver in the LINDO API,
\newblock \emph{Optim. Methods Softw.}, \textbf{24(4-5)} (2009), 657--668.

\bibitem{locatelliSchoenGO}[10.1137/1.9781611972672]
\newblock M. Locatelli and F. Schoen,
\newblock \emph{Global optimization: theory, algorithms and applications},
\newblock SIAM (2013).

\bibitem{luqueAlbaGA}[10.1007/978-3-642-22084-5]
\newblock G. Luque and E. Alba,
\newblock \emph{Parallel genetic algorithms. Theory and real world applications},
\newblock Springer-Verlag, Berlin (2011).

\bibitem{mongeauKarsentyRouze2000}[10.1080/10556780008805783]
\newblock M. Mongeau, H. Karsenty, V. Rouzé and J. B. Hiriart-Urruty,
\newblock Comparison of public-domain software for black box global optimization,
\newblock \emph{Optim. Methods Softw.}, \textbf{13} (2000), 203--226.

\bibitem{mullen2014}[10.18637/jss.v060.i06]
\newblock K. M. Mullen,
\newblock Continuous global optimization in R,
\newblock \emph{J. Stat. Softw.}, \textbf{60(6)} (2014).

\bibitem{pardalosZhigljavskyZilinskas2016}[10.1007/978-3-319-29975-4]
\newblock M. P. Pardalos, A. A. Zhigljavsky and J. Zilinskas,
\newblock \emph{Advances in stochastic and deterministic global optimization},
\newblock Springer (2016).

\bibitem{pinterGO}[10.1007/978-1-4757-2502-5]
\newblock J. D. Pinter,
\newblock \emph{Global optimization in action (Continuous and Lipschitz optimization: algorithms, implementations and applications)},
\newblock Kluwer Academic Publishers, Dordrecht (1996).

\bibitem{pinter2009}
\newblock J. D. Pinter,
\newblock Software development for global optimization,
\newblock \emph{Lectures on Global Optimization. Fields Institute Communications}, \textbf{55} (2009), 183--204.

\bibitem{riosSahinidis2013}[10.1007/s10898-012-9951-y]
\newblock L. M. Rios and N. V. Sahinidis,
\newblock Derivative-free optimization: a review of algorithms and comparison of software implementations,
\newblock \emph{J. Glob. Optim.}, \textbf{56} (2013), 1247--1293.

\bibitem{sahinidis1996}[10.1007/BF00138693]
\newblock N. V. Sahinidis,
\newblock BARON: A general purpose global optimization software package,
\newblock \emph{J. Glob. Optim.}, \textbf{8} (1996), 201--205.

\bibitem{sergeyev1995}[10.1137/0805041]
\newblock Y. D. Sergeyev,
\newblock An information global optimization algorithm with local tuning,
\newblock \emph{SIAM J. Optim.}, \textbf{5(4)} (1995), 858--870.

\bibitem{sergeyev1999}
\newblock Y. D. Sergeyev,
\newblock Multidimensional global optimization using the first derivatives,
\newblock \emph{Comput. Math. Math. Phys.}, \textbf{39(5)} (1999), 743--752,.

\bibitem{sergeyevGrishagin1994}[10.1007/BF02207778]
\newblock Y. D. Sergeyev and V. A. Grishagin,
\newblock A parallel method for finding the global minimum of univariate functions,
\newblock \emph{Journal of Optimization Theory and Applications}, \textbf{80(3)} (1994), 513--536.

\bibitem{sergeyevGrishagin2001}[10.1023/A:1010185125012]
\newblock Y. D. Sergeyev and V. A. Grishagin,
\newblock Parallel asynchronous global search and the nested optimization scheme,
\newblock \emph{J. Comput. Anal. Appl.}, \textbf{3(2)} (2001), 123--145.

\bibitem{sergeyevStronginLera2013}[10.1007/978-1-4614-8042-6]
\newblock Y. D. Sergeyev, R. G. Strongin and D. Lera,
\newblock \emph{Introduction to global optimization exploiting space-filling curves},
\newblock Springer (2013).

\bibitem{Famularo2001}[10.1023/A:1012391611462]
\newblock Ya. D. Sergeyev, D. Famularo and P. Pugliese,
\newblock Index branch-and-bound algorithm for lipschitz univariate global optimization with multiextremal constraints,
\newblock \emph{J. Glob. Optim.}, \textbf{21(3)} (2001), 317--341.

\bibitem{strongin1978}
\newblock R. G. Strongin,
\newblock \emph{Numerical methods in multi-extremal problems (information-statistical algorithms)},
\newblock Moscow: Nauka, In Russian (1978).

\bibitem{strongin1992}[10.1007/BF00122428]
\newblock R. G. Strongin,
\newblock Algorithms for multi-extremal mathematical programming problems employing a set of joint space-filling curves,
\newblock \emph{J. Glob. Optim.}, \textbf{2(4)} (1992), 357--378.

\bibitem{stronginGergelBarkalovParGO}
\newblock R. G. Strongin, V. P. Gergel, V. A. Grishagin and K. A. Barkalov,
\newblock \emph{Parallel computations for global optimization problems},
\newblock Moscow State University (In Russian), Moscow (2013).

\bibitem{strSergGO}[10.1007/978-1-4615-4677-1]
\newblock R.G. Strongin and Ya.D. Sergeyev,
\newblock \emph{Global optimization with non-convex constraints. Sequential and parallel algorithms},
\newblock Kluwer Academic Publishers, Dordrecht (2000).

\bibitem{zilinskTornGO}[10.1007/3-540-50871-6]
\newblock A. Törn and A. Zilinskas,
\newblock \emph{Global optimization},
\newblock Springer (1989).

\bibitem{venkataraman2009}
\newblock P. Venkataraman,
\newblock \emph{Applied optimization with MATLAB programming},
\newblock John Wiley \& Sons (2009).

\bibitem{zhigljavskyRandGO}[10.1007/978-94-011-3436-1]
\newblock A. A. Zhigljavsky,
\newblock \emph{Theory of global random search},
\newblock Kluwer Academic Publishers, Dordrecht (1991).

\end{thebibliography}

\medskip
% The data information below will be filled by AIMS editorial staff
Received xxxx 20xx; revised xxxx 20xx.
\medskip

\end{document}

### 2017-05-25 2017/NACO/paper.tex
Update Globalizer paper

@@
 As it has been shown earlier by many researchers
-(see, for instance, [12], [25], [38], [51]),
+(see, for instance, \cite{floudasPardalosGOState}, \cite{horstTuyGO}, \cite{pinterGO}, \cite{strSergGO}),
 finding the numerical estimate of the global optimum implies constructing a coverage of
@@
 of multidimensional search information. As a result, many optimization algorithms use
-various approaches to the dimensional reduction [38], [46], [48], [50], [51].
+various approaches to the dimensional reduction \cite{pinterGO}, \cite{sergeyevStronginLera2013}, \cite{strongin1978}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
 \par
@@
 onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
-reduction [46], [48], [50], [51].
+reduction \cite{sergeyevStronginLera2013}, \cite{strongin1978}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
 \par
@@
 \label{sec:parallel}
-Let us consider the parallelization methods used widely in the theoretical and
-practical application of parallel computing within the context of global optimization problems:
+Let us consider the parallelization approaches used widely in the theoretical and
+practical applications of parallel computing within the context of global optimization problems:
 \begin{itemize}
-  \item The distribution of the search domain \(D\) among the available computing units
-  (data parallelization scheme). In the case of optimization problems, this approach is
-  insufficient, since in organizing such computations the subdomain, which contain the
-  sought global minimum, will be processed by only one processor, and, therefore,
-  all of the remaining processors would perform the excess computations.
-  \item The parallelization of the computational schemes for optimization algorithms
-  (task parallelization scheme). This approach is also insufficient since the direct
-  computational costs for executing optimization algorithms are relatively low
-  (the majority of computations in the course of a global search are represented by
-  calculations of the optimized function values due to the initial assumption of
-  considerable computational costs for such calculations).
-  \item The parallelization of the computations executed in order to obtain values
-  for the optimized function. This approach can be applied since the most computationally
-  costly part of the global search process will be parallelized. However, this method is not
-  featured by the generality (the development of parallelization methods is to be performed
-  every time from scratch while solving each particular optimization problem).
+  \item The distribution of the search domain \(D\) among the available computational units (data parallelization scheme).
+This approach is insufficient in the case of optimization problems since in such computations the subdomain, which contain the sought global minimum, will only be processed by one processor, and, therefore, all the remaining processors would perform the excess computations.
+  \item The parallelization of the optimization algorithms (task parallelization scheme).
+This approach is also insufficient since the direct computational costs for executing the optimization algorithms are relatively low (the majority of computations are represented by the calculations of the optimized function values due to the initial assumption of considerable computational costs of such calculations).
+  \item The parallelization of the computations executed in order to obtain the values of the optimized function.
+This approach is applicable since the most time-consuming part of the global optimization will be parallelized.
+However, this method is not featured by the generality (the development of parallelization methods is to be performed every time from the very beginning while solving each particular optimization problem).
 \end{itemize}
 \par
-Within the framework of information-statistical theory, a general approach to
-parallelization computations when solving global optimization problems has been
-proposed --- \textit{the parallelism of computations is provided by means of simultaneously
-computing the values of the minimized function \(\varphi(y)\) at several different
-points within the search domain \(D\)} – see, for example, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
-This approach provides parallelizaion for the most costly part of computations in the global search process.
-\par
-The global optimization algorithms implemented in Globalizer will be described step-by-step below.
-In Subsection \ref{subsec:corepar}, the core sequential multidimensional algorithm of global search (MAGS) will be presented.
-In Subsection \ref{subsec:sharedpar}, a parallel generalization of the MAGS algorithm for parallel computations on
-computer systems with shared memory will be described. In Subsection \ref{subsec:distribpar}, the scheme
-for parallel computations by multiprocessor systems with distributed memory will be provided.
-
+Within the framework of the information-statistical theory, a general parallelization
+approach for solving the global optimization problems has been proposed [\cite{stronginGergelBarkalovParGO}, \cite{strSergGO} --- \textit{the
+parallel computations are provided by means of simultaneous computing the values of the
+minimized function \(\varphi(y)\) at several different points within the search domain \(D\)}.
+This approach provides the parallelization for the most time-consuming part of the optimization computations.
+\par
+The global optimization algorithms implemented in Globalizer will be described step-by-step.
+In Subsection \ref{subsec:corepar}, the core serial Multidimensional Algorithm of Global
+Search (MAGS) is presented. In Subsection \ref{subsec:sharedpar}, a parallel generalization
+of the MAGS algorithm for the parallel computations on the multiprocessor computational
+nodes with shared memory is described. In Subsection \ref{subsec:distribpar}, the scheme
+for the parallel computations on the high-performance computational systems with distributed memory is given.
 \subsection{Core Multidimensional Generalized Algorithm of Global Search}
 \label{subsec:corepar}
-The information-statistical theory of global search formulated in \cite{strongin1978}, \cite{strSergGO}
-has served as a basis for the development of a large number of efficient multiextremal optimization methods --- see, for example, \cite{barkalovGergel2014}, \cite{Pizzuti}, \cite{gergel1996}, \cite{gergel1997}, \cite{grishaginStrongin1984}, \cite{sergeyev1995}, \cite{sergeyev1999}, \cite{sergeyevGrishagin1994}, \cite{sergeyevGrishagin2001}, \cite{sergeyevStronginLera2013}, \cite{Famularo2001} etc.
-\par
-Multidimensional Algorithm of Global Search (MAGS) established the basis for the
-methods applied in Globalizer. The general computational scheme of MAGS can be
-presented as follows --- see \cite{strongin1978}, \cite{strSergGO}.
-\par
-Let us introduce a simpler notation for the problem being solved on a computing node
+The information-statistical theory of global optimization formulated in \cite{strongin1978}, \cite{strSergGO} has
+served as a basis for the development of a large number of efficient multiextremal optimization
+methods \cite{barkalovGergel2014}, \cite{Pizzuti}, \cite{gergel1996}, \cite{gergel1997}, \cite{grishaginStrongin1984}, \cite{sergeyev1995}, \cite{sergeyev1999}, \cite{sergeyevGrishagin2001}, \cite{sergeyevStronginLera2013}, \cite{Famularo2001}.
+\par
+The optimization methods applied in Globalizer are based on the MAGS method, which can be presented as follows --- see \cite{strongin1978}, \cite{strSergGO}.
+\par
+Let us introduce a simpler notation for the optimization problem being solved
 \begin{equation}
@@
 The initial iteration of the algorithm is performed at an arbitrary point \(x^1(0,1)\).
-Let us assume further \(k\), \(k>1\), iterations of a global search are to be completed.
+Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
 The selection of the trial point \(k+1\) for the next iteration is performed according to the following rules.
 \par
-\textit{Rule 1}. To renumber the points for the preceding trials by the lower indices in order of increasing value of coordinates
+\textit{Rule 1}. Renumber the points of the preceding trials by the lower indices in order of increasing value of coordinates
 \begin{equation}
@@
 \end{equation}
-The points \(x_0\), \(x_k\) were introduced additionally for convenience in further explanation,
-the values of the minimized function \(z_0\), \(z_k\) at these points are undefined.
-\par
-\textit{Rule 2}. To compute a current estimate of the H{\"o}lder constant \(H\) from (\ref{eq:holder})
+the points \(x_0\), \(x_k\) were introduced additionally for the convenience of further
+explanation, the values of the minimized function \(z_0\), \(z_k\) at these points are undefined.
+\par
+\textit{Rule 2}. Compute current estimate of the H{\"o}lder constant \(H\) from (\ref{eq:holder})
 \begin{equation} \label{step2}
@@
 \end{equation}
-as the maximum of the relative differences of the minimized function \(f(x)\) from (\ref{eq:oneDimFunc})
-on the set of executed trial points \(x_i,1\leq i\leq k\) from (\ref{step1}). Further, \(\rho_i=(x_i-x_{i-1})^\frac{1}{N},1\leq i\leq k+1\).
-The constant \(r\), \(r>1\), is the \textit{parameter} for the algorithm.
-\par
-\textit{Rule 3}. For each interval \((x_{i-1},x_i),1\leq i\leq k+1\), compute the characteristics \(R(i)\) where
+as the maximum of the relative differences of the minimized function values on the
+set of previously executed trial points \(x_i,1\leq i\leq k\) from (\ref{step1}).
+Hereafter, \(\rho_i=(x_i-x_{i-1})^\frac{1}{N},1\leq i\leq k+1\). The
+constant \(r\), \(r>1\), is the reliability parameter of the algorithm.
+\par
+\textit{Rule 3}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq k+1\), where
 \[
@@
 \par
-\textit{Rule 4}. To determine the interval with the maximum characteristic
+\textit{Rule 4}. Determine the interval with the maximum characteristic
 \begin{equation} \label{step4}
@@
 \par
-\textit{Rule 5}. To execute a new trial (computation of the minimized function value \(f(x)\))
-at the point \(x^{k+1}\) located within the interval with the maximum characteristic from (\ref{step4})
+\textit{Rule 5}. Execute a new trial at the point \(x^{k+1}\) located within the interval
+with the maximum characteristic from (\ref{step4})
 \begin{equation} \label{step5}
@@
 \par
-The stop condition by which the trials are terminated, is defined by the inequality
+The stop condition, which terminated the trials, is defined by the inequality
 \begin{equation}
@@
 for the interval with the maximum characteristic from (\ref{step4}) and \(\varepsilon >0\) is the predefined
-accuracy of the problem solution. If the stop condition is not satisfied, the index \(k\) is incremented by unity,
-and a new global search iteration is executed.
+accuracy of the optimization problem solution. If the stop condition is not satisfied,
+the index \(k\) is incremented by 1, and new global optimization iteration is executed.
 \par
 In order to explain the algorithm presented above, let us note the following.
-The characteristics \(R(i), 1\leq i\leq k+1\), calculated according to (\ref{step3})
-could be interpreted as some measures of importance to the intervals with respect to
-location of the global minimum point. Thus, the scheme for selecting the interval
-for executing the next trial described in (\ref{step4}), (\ref{step5}) becomes more clear --- the point of every next
-global search iteration is selected within the interval where the global minimum point can most likely be found.
-\par
-The conditions for the algorithm’s convergence described above were examined, for example, in \cite{strSergGO}.
+The characteristics \(R(i), 1\leq i\leq k+1\), calculated according to (\ref{step3}) could
+be interpreted as some measures of importance of the intervals with respect to the
+expected location of the global minimum point. Thus, the rules (\ref{step4}) and (\ref{step5}) for selecting
+the interval for executing the next trial become more clear --- the point of every next
+global optimization iteration is selected within the interval, where the global minimum
+point can be found most likely.
+\par
+The convergence conditions for the described algorithm are given, for example, in \cite{strSergGO}.
 
@@
 \label{subsec:sharedpar}
-Modern supercomputer computational systems consist of plenty of computer nodes,
-which include several multicore processors. At that, random access memory is
-shared for the computing nodes --- the content of any memory element can be read (written)
-by any computer cores available at any arbitrary moment. In most cases, shared memory is
-uniform --- the time characteristics for accessing memory are the same for all computer
-cores and for all memory elements.
-\par
-The following speculations can form the basis for selecting parallel computation
-organization methods. As was mentioned above when describing the MAGS algorithm,
-the characteristics \(R(i),1\leq i\leq k+1\), calculated according to (\ref{step3})
-can be interpreted as some measures of the importance of intervals with respect to the location of
-the global minimum point. Following this understanding, each successive global search
-iteration is executed within the most important interval with the maximum value of the
-characteristic. So far, in this case it is obvious how to select the other intervals for
-organizing simultaneous computations of the minimized function values at several different
-points within the search domain as well --- these should be the intervals with the next
-magnitudes of characteristics.
-\par
-The computational scheme of the Parallel Multidimensional Algorithm of Global Search for
-computer systems with shared memory (PMAGS) was developed based the approach considered
-above and is practically identical to the MAGS scheme --- the differences consist just in the
-following set of rules --- see \cite{strSergGO}, \cite{stronginGergelBarkalovParGO}.
-\par
-\textit{Rule 4'}. To arrange the characteristics of the intervals obtained according to (\ref{step3}) in decreasing order
+Modern supercomputer computational systems consist of plenty of computational nodes,
+which include several multicore processors. In this case, random access memory is
+shared among the computational nodes --- the content of any memory location can be read
+(written) by any computational core at any arbitrary moment. In most cases, shared memory is
+uniform --- the time characteristics for accessing the memory are the same for all
+computational cores and for all memory locations.
+\par
+The following approach can be applied to provide the parallel computations with the MAGS method.
+As it was mentioned above, the characteristics \(R(i),1\leq i\leq k+1\) calculated
+according to (\ref{step3}) can be interpreted as some measures of importance of the
+intervals with respect to the expected location of the global minimum point.
+Following this understanding, every next global optimization iteration is executed within
+the most important interval with the maximum characteristic. As a result, it becomes
+evident how to select the other intervals for simultaneous computations of the minimized
+function values at several different points within the search domain --- these ones
+should be the intervals with the next magnitudes of characteristics.
+\par
+The computational scheme of the Parallel Multidimensional Algorithm of Global Search (PMAGS)
+for the computational systems with shared memory can be developed based on the approach
+described above. This scheme is almost identical to the MAGS one --- the differences
+consist just in the following set of rules.
+\par
+\textit{Rule 4'}. Arrange the characteristics of the intervals computed according to (\ref{step3}) in the decreasing order
 \begin{equation}
@@
 \end{equation}
-and to select \(p\) intervals with the indices \(t_j,1\leq j\leq p\), with
-the highest values of characteristics (\(p\) is the number of processors/cores used for the parallel computations).
-\par
-\textit{Rule 5'}. To execute new trials (computations of values for the minimized
-function \(f(x)\)) at the points \(x_{k+j},1\leq j\leq p\), located in the
-intervals with the highest characteristics from (\ref{step4par})
+and select \(p\) intervals with the indices \(t_j,1\leq j\leq p\), with the maximum
+values of characteristics (\(p\) is the number of processors/cores used for the parallel computations).
+\par
+\textit{Rule 5'}. Execute new trials at the points \(x_{k+j},1\leq j\leq p\) located
+in the intervals with the maximum characteristics from (\ref{step4par})
 
@@
 \par
-The stop condition for the algorithm (\ref{eq:stop}) should be checked for all
-intervals from (\ref{step4par}), in which the scheduled trials are executed
+The stop condition (\ref{eq:stop_1}) should be checked for all intervals from (\ref{step4par}),
+where the scheduled trials are executed
 \begin{equation}
@@
 \end{equation}
-As before, if the stop condition is not fulfilled, the index \(k\) is
-incremented by \(p\), and a new global search iteration is executed.
-\par
-The conditions for the developed parallel algorithm’s convergence and for the
-non-redundancy of parallel computations have been examined in \cite{strSergGO}.
-Thus, in particular, when the conditions for convergence are satisfied, the parallel
-computations are non-redundant as compared to the sequential method when using up to
-\(2^N\) processors/cores (where \(N\) is the dimensionality of the global optimization problem being solved).
+If the stop condition is satisfied at least for one interval \(t_j, 1\le j\le p\), the calculations
+are terminated. Otherwise, the index \(k\) is incremented by \(p\), and new global optimization iteration is executed.
+\par
+The convergence conditions for the developed parallel algorithm are considered
+in \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}. Thus, in particular, when the convergence conditions are
+satisfied, the parallel computations are non-redundant as compared to the serial method
+when using up to \(2^N\) processors/cores (\(N\) is the dimensionality of the global optimization problem).
 
@@
 \label{subsec:distribpar}
-The next level of parallel computations in high-performance computer systems consists of
-using several computational nodes. For that, each computational node has its own separate
-memory and, in this case, the interaction between different nodes can be provided by
-means of data transfer only via the computer communication network.
-\par
-To organize parallel computations for multiprocessor systems with distributed memory, using a set of mappings
+The next level of parallel computations on the high-performance computational systems
+consists in using several computational nodes. In this case, each computational node has
+its own separate memory and the interaction between different nodes can be provided by
+means of data transfer via the computer communication network.
+\par
+To provide parallel computations on high-performance computational systems with the distributed memory, it is proposed to use a set of mappings
 \begin{equation}
@@
 \end{equation}
-instead of applying a single Peano curve \(y(x)\) has been proposed in \cite{strongin1992}, \cite{strSergGO}.
-In order to build the set \(Y_s(x)\), several different approaches can be applied.
-Thus, for example, in \cite{strongin1992} a scheme has been applied by which each
-mapping \(y_i(x)\) from \(Y_s(x)\) is built as the result of some shift along the main
-diagonal of the hypercube \(D\). This way, the set of constructed Peano curves enables the
-close prototypes \(x'\), \(x''\)  to be obtained from the interval \([0,1]\) for any close
-multidimensional images \(y'\), \(y''\) from \(D\) differing in one coordinate,
-only for some mapping \(y_k(x), 1\leq k\leq s\).
-\par
-Some other methods for constructing multiple mappings have been examined in \cite{stronginGergelBarkalovParGO}.
-\par
-The set of mappings \(Y_s(x)\) from (\ref{eq:maps}), for multidimensional problem (\ref{eq:task})
-generates \(s\) subsidiary information-linked one-dimensional problems of the kind (\ref{eq:oneDimTask}):
+instead of applying a single Peano curve \(y(x)\) --- see, for instance,
+\cite{strongin1992}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
+In order to construct the set \(Y_s(x)\), several different approaches can be applied.
+Thus, for example, in \cite{strongin1992} each mapping \(y^i(x)\) from \(Y_s(x)\) is constructed
+as the result of some shift along the main diagonal of the hypercube \(D\). This way,
+the set of the constructed Peano curves enables the close prototypes \(x'\), \(x''\)
+to be obtained from the interval \([0, 1]\) for any close multidimensional images
+\(y'\), \(y''\) from \(D\) differing in one coordinate only.
+\par
+Some other methods for constructing the multiple mappings were considered in \cite{stronginGergelBarkalovParGO}.
+\par
+The set of mappings \(Y_s(x)\) from (\ref{eq:maps}) generates \(s\) information-linked one-dimensional
+problems of the kind (\ref{eq:oneDimTask}):
 \begin{equation}
@@
 \par
-It is important to note that the family of one-dimensional problems
-\[
-\varphi(y^l(x)),1 \leq l \leq s,
-\]
-obtained as a result of the dimensional reduction is information-linked --- the function
-values computed for any problem \(\varphi(y^l(x))\) from the family (\ref{eq:oneDimProblemSerie}) can be used for all
-of the remaining problems of this family.
-\par
-The information compatibility of the problems from the family (\ref{eq:oneDimProblemSerie}) allows the parallel
-computations to be organized in an obvious enough way. Thus, each particular problem can
-be solved by a separate processor in the computing system; the exchange of search information
-between the processors should be performed during the course of the computations.
-As a result of using such a parallelization method, a unified approach for organizing parallel
-computations for a multiprocessor computer systems with distributed, as well as shared
-memory can be developed. This method consists of the following.
+It is important to note that the family of the one-dimensional problems
+\(\varphi(y^l(x)),1 \leq l \leq s\) obtained as a result of the dimensionality
+reduction is an information-linked one --- the function values computed for any
+problem \(\varphi(y^l(x))\) from the family (\ref{eq:oneDimProblemSerie}) can be used for all rest problems of this family.
+\par
+The information compatibility of the optimization problems from the family (\ref{eq:oneDimProblemSerie})
+allows executing the parallel computations in the following way. Each particular
+problem can be solved by a separate processor of the computational system; the exchange
+of search information between the processors should be performed during the computations.
+As a result, a unified approach for the parallel computations on the high-performance
+computational systems with distributed and shared memory can be developed. This method consists in the following.
 \begin{enumerate}
-  \item The family of reduced one-dimensional information-linked problems (\ref{eq:oneDimProblemSerie}) is
-  distributed among the computational nodes of a multiprocessor system. One problem,
-  as well as several others from within the family, can be allocated to each particular computational node.
-  \item The Parallel Multidimensional Algorithm of Global Search (PMAGS, see Subsection \ref{subsec:sharedpar})
-  is applied to solve the allocated problems from the family (\ref{eq:oneDimProblemSerie})
-  at each computational node, supplemented by the following rules of information interaction.
+  \item The family of the reduced one-dimensional information-linked optimization
+  problems (\ref{eq:oneDimProblemSerie}) is distributed among the computational nodes of the multi-node computational system.
+  \item Each optimization problem from the family (\ref{eq:oneDimProblemSerie}) is solved on the
+  corresponding computational node using the PMAGS method (described above in Subsection \ref{subsec:sharedpar})
+  supplemented by the following rules of the information interaction.
   \begin{enumerate}
-    \item Prior to beginning a new trial at any point \(x'\in [0,1]\) for any problem \(\varphi(y_l(x)),1\leq l\leq s\),
-    the following should be performed:
+    \item Prior to beginning new trial at any point \(x'\in [0,1]\) for any problem \(\varphi(y^l(x)),1\leq l\leq s\), the following should be performed:
     \begin{itemize}
-      \item compute the image \(y'\in D\) for the point \(x'\in [0, 1]\) according to the mapping \(y_l(x)\),
-      \item determine  the prototypes \(x_i',1\leq i\leq s\), for each
-      problem from the family (\ref{eq:oneDimProblemSerie}),
-      \item send the prototypes \(x_i',1\leq i\leq s\) to all computational
-      nodes employed in order to exclude the repeated selection of the intervals in
-      which the prototypes fall, for using them to determine the points for new trials.
-      To organize the data transfer, a queue can be formed at each computational
-      node for sending the trial points and receiving the minimized function values at these points.
+      \item compute the image \(y'\in D\) for the point \(x'\in [0, 1]\) according to the mapping \(y^l(x)\),
+      \item determine  the preimages\footnote{All preimages have the same
+      image \(y'\); the preimages are calculated by the inverse mappings \((y^i)^{-1}\), \(1\le i\le s\);
+      the number of the preimages coincides with the number of the mappings.}
+      \begin{displaymath}
+        x_i':y'=y^i(x_i'),1\le i\le s,
+      \end{displaymath}
+      in accordance with the set of mappings \(y^l(x), 1\le l\le s\) for each problem from the family (\ref{eq:oneDimProblemSerie}),
+      \item send the prototypes \(x_i',1\leq i\leq s\) to all computational nodes in order
+      to exclude the repeated selection of the intervals, which the prototypes fall in,
+      for using these ones to determine the points of new trials. To perform the data transfer,
+      a queue can be created at each computational node for sending the trial points and
+      receiving the minimized function values at these points.
     \end{itemize}
-    \item After completing any trial for any problem \(\varphi(y_l(x)),1\leq l\leq s\), at any point
-    \(x'\in[0,1]\), it is necessary:
+    \item After completing any trial for any problem \(\varphi(y_l(x)),1\leq l\leq s\),
+    at any point \(x'\in[0,1]\), it is necessary:
     \begin{itemize}
-      \item to determine all prototypes \(x_i',1\leq i\leq s\), for the point of the completed trial
-      for each problem from the family (\ref{eq:oneDimProblemSerie}),
-      \item to send the prototypes \(x_i',1\leq i\leq s\), and the result of the
-      trial \(z'=\varphi(y_l(x'))\) to all computational nodes employed to include the
-      obtained data into the search information processed within the rules of the parallel global search algorithm.
+      \item to determine all prototypes \(x_i',1\leq i\leq s\) for the point of the
+      completed trial for each problem from the family (\ref{eq:oneDimProblemSerie}),
+      \item to send the prototypes \(x_i',1\leq i\leq s\) and the result of the trial
+      \(z'=\varphi(y_l(x'))\) to all computational nodes to include the obtained data into
+      the search information processed according to the rules of the parallel global optimization algorithm.
     \end{itemize}
-    \item Prior to starting the next global search iteration, the algorithm should
-    check the queue of the received data; if there are some data in the queue,
-    they should be included in the search information.
+    \item Prior to starting the next global optimization iteration, the algorithm
+    should check the queue of the received data; if there are any data in the queue,
+    these should be included into the search information. The MAGS algorithm uses these
+    accumulated data for the adaptive computing of the optimization iteration points (see Section \ref{subsec:corepar}).
 \end{enumerate}
@@
 \par
-The possibility of asynchronous data transfer (the computational nodes process received
-data only upon receipt) is a meaningful feature of such a scheme for organizing parallel computations.
-\par
-In addition, there is no single managing node within this scheme. The number of
-computational nodes can vary during the course of the global search, and excluding any
-node does not result in the loss of the sought global minimum of the minimized multiextremal function.
+The possibility of asynchronous data transfer (the computational nodes process the received
+data upon receipt only) is an important feature of the proposed parallel computation scheme.
+In addition, there is no any selected managing node within this scheme. The number of
+computational nodes can vary during the global optimization, and excluding any node does not result
+in the loss of the sought global minimum of the minimized multiextremal function.
 \par
 This computational scheme generates the Generalized Parallel Multidimensional Algorithm of
-Global Search (GPMAGS) for high-performance computing systems, which may include many
-multiprocessor (multicore) computational nodes with distributed memory as well as
-accelerators for computations based on Intel Xeon Phi processors and on general purpose graphic processors.
-\par
-Additional information on organizing such parallel computation schemes is provided in \cite{gergelSidorov2015}.
+Global Search (GPMAGS) for high-performance computational systems, which may include many
+multiprocessor (multicore) computational nodes with distributed memory as well as the
+computational accelerators based on Intel Xeon Phi processors and on the general purpose graphic processors (GPUs).
+\par
+Additional information on such parallel computation schemes can be found in \cite{gergelSidorov2015}.
 
@@
 \label{sec:arch}
-As previously mentioned in Section \ref{sec:intro}, the development of software systems
-for global optimization is complex professional work. The optimization systems should
-include advanced optimization methods, provide for the accumulation and efficient
-processing of search information, include various forms of visualization for the results,
-provide tools for interacting with users, etc. All the features pointed out above
-become much more complicated when organizing parallel computations using high-performance
-multiprocessor systems of varying architectures.
-\par
-The Globalizer considered in this paper expands the family of global optimization
-software systems successively developed by the authors during the past several years.
-One of the first developments was the SYMOP multiextremal optimization system \cite{gergel1993},
-which has been successfully applied for solving many optimization problems. A special
-place is occupied by the ExaMin system (see, for example, \cite{barkalovGergel2015}),
-which was developed and used extensively to investigate the application of novel parallel
-algorithms to solve global optimization problems using high-performance multiprocessor computing systems.
+Globalizer expands the family of the global optimization software systems developed
+by the authors during the past several years. One of the first developments was the
+SYMOP multiextremal optimization system \cite{gergel1993}, which has been applied for solving
+many optimization problems. The ExaMin system \cite{barkalovGergel2015}, [3], [4], [16] was developed
+and used to investigate various parallel algorithms for solving the global
+optimization problems on the high-performance computational systems.
 
@@
     \includegraphics[width=0.75\textwidth]{pictures/globalizerScheme.eps}
-    \caption{The Globalizer system architecture (Blocks 1-2, 5-7 have been implemented; Blocks 3-4 and 8-11 are under development)}
+    \caption{The architecture of the Globalizer system}
     \label{fig:globalizerScheme}
@@
 \par
-The Globalizer system architecture is presented in Figure \ref{fig:globalizerScheme}. The structural components are:
+The architecture of the Globalizer system is presented in Figure \ref{fig:globalizerScheme}.
+The system components are:
 \begin{itemize}
-  \item Block 0 consists of the procedures for computing the function values (criteria
-  and constraints) for the optimization problem being solved; this is an external
-  block with respect to the Globalizer system and is provided by the user per a predefined interface.
+  \item Block 0 consists of the procedures for computing the function values
+  (criteria and constraints) of the optimization problem being solved; it is an external
+  block with a predefined interface with respect to the Globalizer system and is provided by the user.
   \item Blocks 1-4 form the optimization subsystem and solve the global optimization
-  problems (Block 1), nonlinear programming (Block 2), multicriterial optimization (Block 3),
-  and general decision making problems (Block 4). It is worth noting the sequential
-  scheme of interaction between these components --- the decision making problems are
-  solved using the multicriterial optimization block, which, in turn, uses the nonlinear programming block, etc.
-  \item Block 5 is a subsystem for accumulating and processing the search information;
-  this is one of the basic subsystems --- the amount of search information for complex
-  optimization problems may appear to be quite large on the one hand, but, on the other
-  hand, the efficiency of the global optimization methods depends to a great extent on how
-  completely all of the available search data is utilized.
-  \item Block 6 contains the dimensional reduction procedures based on Peano evolvents;
-  the optimization blocks solve the reduced (one-dimensional) optimization problems;
-  Block 6 provides interaction between the optimization blocks and the initial
-  multidimensional optimization problem.
-  \item Block 7 organizes the choice of parallel computation schemes in the Globalizer
-  system subject to the computing system architecture employed (the numbers of cores in the
-  processors, the availability of shared and distributed memory, the availability of
-  accelerators for computations, etc.) and the global optimization methods applied.
-  \item Block 8 is responsible for managing the parallel processes when performing the
-  global search (determining the optimal configuration of parallel processes, distributing
-  the processes between computer elements, balancing computational loads, etc.).
-  \item Block 9 is a common managing subsystem, which fully controls the whole
-  computational process when solving global optimization problems.
-  \item Block 10 is responsible for organizing the dialog interaction with users for
-  stating the optimization problem, adjusting system parameters (if necessary), and
-  visualizing and presenting the global search results.
-  \item Block 11 is a set of tools for visualizing and presenting the global search results;
-  the availability of tools for visually presenting the computational results enables
-  the user to provide efficient control over the global optimization process.
+  problems (Block 1), nonlinear programming ones (Block 2), multicriterial optimization problems
+  (Block 3), and general decision making problems (Block 4). The interaction between these
+  components is organized in a hierarchical order --- the decision making problems are
+  solved using the multicriterial optimization block, which, in turn, uses the nonlinear programming block, and so on.
+  \item Block 5 is a subsystem for accumulating and processing the search information; it is one of the main subsystems – the size of search information may appear to be quite large and the efficiency of the global optimization methods depends on how completely the available search data are utilized.
+  \item Block 6 contains the dimensionality reduction procedures based on the Peano space-filling curves; the reduced (one-dimensional) optimization problems are solved by the optimization blocks.
+This block also provides the interaction between the optimization blocks and the multidimensional optimization problem being solved.
+  \item Block 7 provides tuning the parallel computation schemes in the Globalizer system subject to the employed computational system architecture (the number of the computational cores, the availability of shared and/or distributed memory, the availability of computational accelerators, etc.) and the applied global optimization methods.
+  \item Block 8 is responsible for managing the parallel processes when performing the global optimization (determining the optimal configuration of parallel processes, distributing the processes between the computational unit, balancing the computational loads, etc.).
+  \item Block 9 is a managing subsystem, which controls the parallel computations when solving the global optimization problems.
+  \item Block 10 provides the dialog interaction with the users for stating the optimization problem, adjusting the system parameters (if necessary), and visualizing and presenting the global optimization results.
+  \item Block 11 is a set of tools for visualizing and presenting the global optimization results; the availability of the tools for visual demonstration of the computational results enables the user to perform an efficient control over the global optimization process.
 \end{itemize}
 \par
-The Globalizer system comes in two different options:
+The Globalizer system comes in two different versions:
 \begin{itemize}
-  \item Globalizer Pro is a research version of the system providing a full set of
-  capabilities that are oriented toward solving of the most complicated global optimization problems.
-  \item Globalizer Lite is a limited version of the Globalizer system accessible for free for
-  solving global optimization problems with a medium level of difficulty, and is accessible free of charge.
+  \item Globalizer Pro is a research version for making various large-scale numerical experiments on the high-performance computational systems to investigate the new parallel global optimization methods being developed.
+  \item Globalizer Lite is a limited version for solving the global optimization problems with a medium complexity level, and is accessible free of charge.
 \end{itemize}
@@
 \label{sec:experiments}
-In order to confirm the operability of the Globalizer system, results from two
-series of computational experiments are presented in this section.
-\par
-The computational experiments were conducted using the Lobachevsky supercomputer at the
-State University of Nizhny Novgorod (operation system --- CentOS 6.4, management system --- SLURM).
-Each supercomputer node had 2 Intel Sandy Bridge E5-2660 processors, 2.2 GHz, with 64 GB RAM.
-Each processor had 8 cores (i.e. a total of 16 CPU cores were available at each node).
-In order to generate the executable program code, the Intel C++ 14.0.2 compiler was used.
-\par
-A number of nodes in the computer system utilized Intel Xeon Phi 5110P processors,
-each containing 60 cores (a total of 240 threads). In addition, the supercomputer also
-included computer nodes equipped with two GPU NVIDIA Kepler K20X, each providing 14
-stream multiprocessors (a total of 2688 CUDA-cores). In order to generate the executable
-code, the CUDA Toolkit 6.0 was used.
-\par
-The problems generated by the GKLS-generator \cite{gavianoKvasovLeraSergeev2003}
-were selected for the test problems. This generator allowed for obtaining the multiextremal
-optimization problems with \textit{a priori} known properties: the number of local minima,
-the sizes of their attractors, the global minimum point, the function value at this point, etc.
-In order to simulate the computation costs inherent in the applied optimization problems (see, for example, \cite{Barkalov2013}, \cite{gergel2013}, \cite{gergel2015}), in all of the experiments computing the objective function was complicated by auxiliary computations, not altering the form of the function and the position of its minima.
-\par
-The computational experiments considered below were first carried out using the ExaMin system (see \cite{barkalovGergel2015}, \cite{barkalovGergelLebedev2015}, \cite{barkalovGergelLebedevSysoev2015}, \cite{gergelLebedev2015}) and then were reproduced using the Globalizer system.
+The numerical experiments were carried out using the Lobachevsky supercomputer at
+the State University of Nizhni Novgorod (the operation system – CentOS 6.4, the
+supercomputer management system – SLURM). Each supercomputer node had 2 Intel Sandy Bridge
+E5-2660 processors, 2.2 GHz, with 64 GB RAM. Each processor had 8 cores (i. e. total
+16 CPU cores were available at each node). In order to generate the executable program code, the Intel C++ 14.0.2 compiler was used.
+\par
+A number of computational nodes had Intel Xeon Phi 5110P processors, each containing 60 cores
+(total 240 threads). In addition, the supercomputer included the computational nodes equipped
+with two NVIDIA Kepler K20X GPUs, each providing 14 stream multiprocessors (total 2688 CUDA-cores). In order to generate the executable code, the CUDA Toolkit 6.0 was used.
+\par
+In the performed experiments the optimization problems generated by the GKLS-generator
+\cite{gavianoKvasovLeraSergeev2003} were used. This generator allows obtaining the multiextremal optimization
+problems with \textit{a priori} known properties: the number of local minima, the sizes of the attractors of these ones, the global minimum point, the function value at this point, etc.
+In order to simulate the computational costs inherent to the applied optimization problems, in all experiments computing the objective function was intentionally complicated by the auxiliary computations not altering the form of the function and the position of its minima.
+\par
+The numerical experiments were carried out first using the ExaMin system and then were reproduced using the Globalizer system.
+In all the executed experiments, 100 global optimization problems generated randomly by the GKLS-generator were solved.
+The results of experiments were averaged over the number of the solved optimization problems.
 
 \subsection{Comparison with over methods}
-
-The results of numerical comparison of three sequential algorithms -- DIRECT \cite{Jones}, DIRECT\textit{l} \cite{Gablonsky} and core global search algorithm (GSA) from section \ref{subsec:corepar} -- are provided below (results of the work of the first two algorithms are given in paper \cite{gavianoKvasovLeraSergeev2003}).
-Numerical comparison was carried out on function classes \textit{Simple} and \textit{Hard} of dimension 4 and 5 from \cite{gavianoKvasovLeraSergeev2003} since solving a problems of dimension 2 and 3 demands a small number of iterations.
-Global minimizer $y^\ast$ was considered to be found, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^\ast\right\|\leq\delta$.
-The size of the vicinity was selected (according to \cite{gavianoKvasovLeraSergeev2003}) as $\delta = \left\|b-a\right\|\sqrt[N]{\Delta}$, $N$ -- problem dimension, $a$ and $b$ -- borders of search domain $D$, parameter $\Delta=10^{-6}$ at $N=4$ and $\Delta=10^{-7}$ at $N=5$.
-When using the GSA for the class \textit{Simple} parameter $r=4.5$ was selected, for the class \textit{Hard} $r=5.6$ was selected; evolvent construction parameter was fixed as $m=10$.
-The maximum allowable number of iterations was $K_{max} = 10^6$.
-
-The average number of iterations $k_{av}$ performed by the method for solving a series of problems from these classes is shown in table \ref{tab:3}.
-Symbol ''$>$''  reflects a situation in which not all problems of the class were solved by the method.
-This means that the algorithm was stopped as the maximum allowable number of iterations $K_{max}$ was achieved.
-In this case, value $K_{max}=10^6$ was used for calculating the average value of the number of iterations $k_{av}$ that corresponds to the lower estimate of this average value.
-The number of unsolved problems is specified in brackets.
+The results of numerical comparison of three sequential algorithms --- DIRECT
+\cite{Jones}, DIRECT\textit{l} \cite{Gablonsky}, and MAGS from Section \ref{subsec:corepar} --- are presented below.
+
+The numerical comparison was carried out by using the function classes Simple and Hard
+of the dimensions 4 and 5. The results of numerical experiments for DIRECT and
+DIRECT\textit{l} are taken from \cite{sergeyevKvasov2006}. According to \cite{sergeyevKvasov2006}, the global minimizer \(y^*\)
+was considered to be found if the algorithm generated a trial point \(y^k\) inside
+a hyperinterval with a vertex \(y^*\) and
+\begin{displaymath}
+    |y^k(i)-y^*(i)|\le \sqrt[N]{\Delta}(b(i)-a(i)), i\le i\le N
+\end{displaymath}
+where \(N\) is the problem dimensionality, a and b are the borders of the search domain \(D\)
+\footnote{In the case of MAGS, the stop condition was different slightly --- the
+global minimizer \(y^*\) was considered to be found, if MAGS generates a trial
+point \(y^k\)  in the \(\delta\)-vicinity of the global minimum, i. e. \(||y^k-y^*||\le\delta\).
+The size of the vicinity was selected  as \(\delta=\left\|b-a\right\|\sqrt[N]{\Delta}\).}.
+
+The values of the parameter \(\Delta = 10^{-6}\) at \(N = 4\) and \(\Delta = 10^{-7}\)
+at \(N = 5\) were used. When using MAGS, the reliability parameter was selected as
+follows: \(r = 4.5\) for the Simple class and \(r = 5.6\) for the Hard class.
+The density level parameter of the Peano curve approximation was fixed as \(m = 10\).
+
+The averaged numbers of iterations $k_{av}$ executed by the global optimization methods
+for solving the optimization problems are shown in Table \ref{tab:3}. The symbol ''$>$''
+denotes the situation, when not all problems of the class are solved by the optimization
+method. This means that the algorithm was stopped as the maximum allowed number of
+iterations $K_{max}$ was achieved. In this case, the value $K_{max}= 10^6$  was used to
+calculate the averaged number of iterations kav that corresponds to the lower estimate
+of this averaged value. The number of the unsolved problems is specified in the brackets.
 
 \begin{table}
-  \caption{Average number of iterations $k_{av}$}
+  \caption{Averaged number of executed iterations of the compared global optimization methods}
   \label{tab:3}
@@
 
-Table \ref{tab:3} shows that GSA outperforms DIRECT and DIRECT\textit{l} methods on all classes of problems in terms of average number of iterations.
-In class 5-\textit{Hard} all three methods failed to solve all problems: DIRECT did not solve 16 problems, DIRECT\textit{l} and GSA, 4 problems.
+Table \ref{tab:3} demonstrates that MAGS overcomes the DIRECT and DIRECT\textit{l} methods
+on all classes of problems in terms of the averaged number of iterations.
+In the 5-Hard class, all three methods failed to solve some problems: DIRECT did not
+solve 16 problems, DIRECT\textit{l} and MAGS --- 4 problems each.
 
 \subsection{Parallel Computations Using Intel Xeon Phi processors}
-The results of the numerical experiments with GPMAGS on an Intel Xeon Phi are provided in Table \ref{tab:exResults}.
-The computations were performed using the \textit{Simple} and \textit{Hard} function classes with the
-dimensions equal to 4 and 5. The parameters \(\Delta=10^{-6}\) at \(N=4\) and \(\Delta=10^{-6}\) at \(N=5\)
-were used. When using the GPMAGS method, the parameter \(r=4.5\) was selected for the
-\textit{Simple} class, and \(r=5.6\) was selected for the \textit{Hard} class.
-\par
-In the first series of experiments, serial computations using MAGS were executed.
-The average number of iterations performed by the method for solving a series of
-problems for each of these classes is shown in row I. The symbol ``\(>\)'' reflects the
-situation where not all problems of a given class were solved by a given method.
-It means that the algorithm was stopped once the maximum allowable number of
-iterations \(K_{max}\) was achieved. In this case, the \(K_{max}\) value was used for
-calculating the average number of iterations corresponding to the lower estimate of
-this average value. The number of unsolved problems is specified in brackets.
-\par
-In the second series of experiments, parallel computations were executed on a CPU.
-The relative ``speedup'' in iterations achieved is shown in row II; the speedup of
-parallel computations was measured in relation to the sequential computations (\(p=1\)).
+As in the previous case, the numerical experiments were performed by using the Simple
+and Hard function classes with the dimensions equal to 4 and 5. The values of the
+parameter \(\Delta=10^{-6}\) at \(N=5\) and \(\Delta=10^{-7}\) at \(N=5\) were used.
+When using the GPMAGS method, the values of the reliability parameter \(r=4.5\) was
+selected for the Simple class and \(r=5.6\) was selected for the Hard class.
+
+The results of the numerical experiments with GPMAGS with Intel Xeon Phi processors are presented in Tables \ref{tab:exResults} and \ref{tab:exResults2}.
+
+In the first series of experiments, the serial computations using MAGS were executed.
+The average numbers of iterations performed by the method for solving the series of optimization problems
+for each class are shown in row I.
+
+In the second series of experiments, the parallel computations were executed using the CPUs.
+The achieved relative speedup in iterations is shown in row II; the speedup of the
+parallel computations is shown relative to the serial computations \((p=1)\).
+
+The final series of experiments was executed using Xeon Phi processors. The results
+of these experiments are shown in row III. In this case, the speedup factors were
+calculated relative to the results of the PMAGS method executed on the CPU using eight cores \((p=8)\).
 
@@
   \label{tab:exResults}
-  \caption{Results of numerical experiments}
+  \caption{Averaged numbers of iterations executed by GPMAGS for solving the test optimization problems}
   \begin{tabular}{cccccccc}
@@
     \begin{center}
-    \textbf{Serial computations.}\\ \textit{Average number} \\ \textit{of iterations.}
+    \textbf{Seriquential trial computations}
     \end{center}		}
@@
 II  & \textbf{Parallel computations}  %\multirow{3}{*}{}
-  & 2 & 2.51 & 2.26 & & 1.19 & 1.36 \\
-& \textbf{of CPU}. & 4 & 5.04 & 4.23 & & 3.06 & 2.86 \\
-& \textit{Speedup} & 8 & 8.58 & 8.79 & & 4.22 & 6.56 \\
+  & 2 & 4762 & 11178 & & 13378 & 109075 \\
+& on CPU & 4 & 2372 & 5972 & & 5203 & 51868 \\
+&  & 8 & 1393 & 2874 & & 3773 & 51868 \\
     \noalign{\smallskip}\hline	\noalign{\smallskip}
 III & \textbf{Parallel computations} %\multirow{3}{*}{}
-  & 60  & 8.13 & 7.32 & & 9.87 & 6.55  \\
-&  o\textbf{f Xeon Phi}. & 120 & 16.33 & 15.82 & & 15.15 & 17.31 \\
-& \textit{Speedup} & 240 & 33.07 & 27.79 & & 38.80 & 59.31 \\
+  & 60  & 171 & 393 & & 382 & 3452  \\
+& on Xeon Phi  & 120 & 85 & 182 & & 249 & 1306 \\
+&  & 240 & 42 & 103 & & 97 & 381 \\
 
@@
 
-\par
-The final series of experiments was executed using a Xeon Phi. The results of these
-computations are shown in row III; in this case, the speedup factor is calculated in
-relation to the PMAGS results on a CPU using eight cores (\(p=8\)). It can be noted that
-relatively high speedup was achieved. For example, solving a problem from the 5-\textit{Hard} class
-required on average only 633 parallel iterations on the Xeon Phi, whereas the number of
-iterations was more than 37 thousand when using all of the CPU’s computing cores.
+\begin{table}
+  \centering
+  \label{tab:exResults2}
+  \caption{Speedup of parallel computations executed by GMAGS}
+  \begin{tabular}{cccccccc}
+    \cline{3-8}\noalign{\smallskip}
+    \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}   \\
+    \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
+    \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
+    \noalign{\smallskip}\hline
+    I &
+    \parbox{0.25\textwidth}{
+    \begin{center}
+    \textbf{Serquential trial computations.}\\ \textit{Average number} \\ \textit{of iterations.}
+    \end{center}		}
+      & 1 & 11953 & 25263 & & 15920 & \(>\)148342 (4)  \\
+    \hline \noalign{\smallskip}
+II  & \textbf{Parallel computations}  %\multirow{3}{*}{}
+  & 2 & 2.52 & 2.32 & & 1.21 & 1.41 \\
+& of CPU. & 4 & 5.05 & 4.24 & & 3.13 & 2.92 \\
+& \textit{Speedup} & 8 & 8.68 & 8.88 & & 4.24 & 6.66 \\
+    \noalign{\smallskip}\hline	\noalign{\smallskip}
+III & \textbf{Parallel computations} %\multirow{3}{*}{}
+  & 60  & 8.18 & 7.37 & & 9.99 & 6.66  \\
+&  of Xeon Phi. & 120 & 16.316 & 15.815 & & 15.215 & 17.317 \\
+& \textit{Speedup} & 240 & 33.133 & 27.827 & & 38.838 & 59.359 \\
+
+    \noalign{\smallskip}\hline
+  \end{tabular}
+\end{table}
 
 \subsection{Parallel Computations Using General Purpose Graphic Processors}
-In the first series of experiments, only one graphic accelerator was used for solving
-100 6-diminsional problems. The method’s reliability parameter \(r=4.5\) was selected.
-\par
-The average problem solving time using a GPU was 10.78 sec. whereas the averaged problem
-solving time using all of the node’s 16 CPU cores was 53.8 sec.; an almost 5-fold improvement was observed.
-\par
-A larger scale computational experiment was carried out for the problems with
-dimensions of \(N=8\) and \(N=10\): a total of 12 nodes within a cluster with 36
-graphic accelerators (three accelerators per node) were employed, therefore, a total
-of 96768 CUDA-cores were employed.
-\par
-The average time to solve the 8-dimensional problem was 405.6 seconds, 5.9 times
-faster than the CPU version of the algorithm. The average time to solve the
-10-dimensional problem was 2,055.8 sec. The CPU was not used to solve the 10-dimensional
-problem because of the high degree of complexity in the computations.
+In the first series of experiments, single graphic accelerator was used for solving 100
+6-dimensional optimization problems. The reliability parameter \(r=4.5\) was selected.
+\par
+The averaged execution time using GPU was 10.78 sec whereas the averaged execution
+time using all 16 CPU cores of the computational node was 53.8 sec; almost 5-fold improvement was observed.
+\par
+A larger scale computational experiment was carried out for the global optimization
+problems with the dimensions of \(N = 8\) and \(N = 10\): total 12 nodes with 36 graphic
+accelerators (three accelerators per the computational node) were employed (total 96,768 CUDA-cores were used).
+\par
+The averaged execution time using the GPUs to solve the 8-dimensional problem was
+405,6 sec that was 5.9 times less than the execution time for CPU. The averaged time
+of solving the 10-dimensional problem was 2,055.8 sec. (in this case, the execution time for CPU
+was not calculated because of the high complexity level of the required computations).
 
@@
 
-\bibitem{sergeyevGrishagin1994}[10.1007/BF02207778]
-\newblock Y. D. Sergeyev and V. A. Grishagin,
-\newblock A parallel method for finding the global minimum of univariate functions,
-\newblock \emph{Journal of Optimization Theory and Applications}, \textbf{80(3)} (1994), 513--536.
+\bibitem{sergeyevKvasov2006}[	arXiv:1103.2056]
+\newblock 44. Ya.D. Sergeyev, D.E. Kvasov,
+\newblock Global search based on efficient diagonal partitions and a set of Lipschitz constants,
+\newblock \emph{SIAM Journal on Optimization}, \textbf{16(3)} (2006), 910--937.
 

### 2017-05-26 2017/NACO/paper.tex
Fix two references in the Globalizer paper

@@
 SYMOP multiextremal optimization system \cite{gergel1993}, which has been applied for solving
-many optimization problems. The ExaMin system \cite{barkalovGergel2015}, [3], [4], [16] was developed
+many optimization problems. The ExaMin system \cite{barkalovGergel2015}, \cite{barkalovGergelLebedev2015},
+\cite{barkalovGergelLebedevSysoev2015}, \cite{gergelLebedev2015} was developed
 and used to investigate various parallel algorithms for solving the global

### 2017-06-08 2017/NACO/paper.tex
Исправлены замечания Гергеля

@@
  \keywords{Global optimization, information-statistical theory, parallel computing, high-performance
-computer systems,  supercomputer technologies.}
+computational systems, supercomputing technologies.}
 
@@
  \centerline{Lobachevsky State University of Nizhny Novgorod}
-   \centerline{23, Gagarin Avenue}
-   \centerline{Nizhny Novgorod, Russia}
+   \centerline{23, Gagarin Avenue, Nizhni Novgorod, Russia}
 }
@@
 in both theory and practice of optimal decision making. In these kinds of problems,
-the criterion to be optimized has several local optima within the search domain,
-which have different values. The existence of several local optima essentially makes
+the optimization criterion can have several local optima within the search domain. 
+The existence of several local optima essentially makes
 finding the global optimum difficult essentially, since it requires examining the whole feasible
@@
 a good professional basis. In general, the global optimization problems can be solved at
-a reasonable time by employing parallel computations on modern supercomputer systems only.
-\par
-The general state of the art in the field of global optimization has been presented in a
-number of key monographs  \cite{floudasPardGO}, \cite{horstTuyGO}, \cite{locatelliSchoenGO}, \cite{pinterGO}, \cite{strSergGO}, \cite{zilinskTornGO}, \cite{zhigljavskyRandGO}, etc.
+a reasonable time by employing parallel computations on modern supercomputing systems only.
+\par
+The general state of the art in the field of global optimization is presented in a
+number of key monographs  \cite{floudasPardGO}, \cite{horstTuyGO}, \cite{locatelliSchoenGO}, \cite{pinterGO}, \cite{strSergGO}, \cite{zilinskTornGO}, \cite{zhigljavskyRandGO}.
 The development of optimization methods, which use the high-performance computational
@@
 \item GlobSol \cite{kearfott2009} is oriented on solving the global optimization
-problems as well as of the systems of the nonlinear equations. The system includes
+problems as well as the systems of the nonlinear equations. The system includes
 the interval methods based on the branch and bound approach. There are some extensions
@@
 The system is widely used to solve the applied problems in various application fields.
-There is a version of the system for exciting on the parallel computational systems.
+There is a version of the system for executing on the parallel computational systems.
 The system is a commercial product, but it is available for the trial usage.
@@
 \label{sec:problem}
-In this paper, the core class\footnote{In general, the Globalizer can be applied for solving
-multicriterial multiextremal multidimensional optimization problems with nonlinear constraints.}
-of optimization problems, which can be solved using
+In this paper, the core class of optimization problems, which can be solved using
 Globalizer, is formulated. This class involves the multidimensional global
@@
 \begin{array}{cr}\\
-  \varphi(y^*)=\min\{\varphi(y):y\in D\} \\
+  \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
   D=\{y\in \mathbf{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
@@
 \end{equation}
-It is supposed, that the objective function \(\varphi(y)\) satisfies the Lipschitz condition
+with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function \(\varphi(y)\) satisfies the Lipschitz condition
 \begin{equation}
@@
 of such a non-uniform coverage could be provided in an adaptive way, when the selection
-of the next trial points is determined by using the search information ((the preceding
+of the next trial points is determined by using the search information (the preceding
 trial points and the values of the minimized function at these points) obtained in the course
 of computations. This necessary condition complicates considerably the computational schemes
-for global optimization methods since it implies a complex analysis of a large amount
+of global optimization methods since it implies a complex analysis of a large amount
 of multidimensional search information. As a result, many optimization algorithms use
@@
 \label{eq:oneDimTask}
-\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}
+\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
 \end{equation}
@@
 \label{eq:holder}
-|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
-\end{equation}
-where the constant H is defined by the relation \(H=4L\sqrt{N}\), \(L\) is the Lipschitz
+|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
+\end{equation}
+where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
 constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem (\ref{eq:task}).
@@
 The algorithms for the numerical construction of the Peano curve approximations are
-presented in \cite{strSergGO}. As an illustration, an approximation of the Peano curve
+given in \cite{strSergGO}. As an illustration, an approximation of the Peano curve
 for the third density level is shown in Figure \ref{fig:peanoC}. Figure \ref{fig:peanoC}
@@
     \includegraphics[width=0.85\textwidth]{pictures/peanoC.eps}
-    \caption{A Peano curve approximation for the third level}
+    \caption{A Peano curve approximation for the third density}
     \label{fig:peanoC}
@@
   function \(\varphi(y(x))\) from (\ref{eq:oneDimTask}),
-  \item After determining the next trial point \(x\), a multidimensional image \(y\) in the mapping \(y(x)\) is calculated,
-  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the multidimensional point \(y\in D\),
+  \item After determining the next trial point \(x\), a multidimensional image \(y\) is calculated by using the
+mapping \(y(x)\),
+  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the point \(y\in D\),
   \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
@@
 Within the framework of the information-statistical theory, a general parallelization
-approach for solving the global optimization problems has been proposed [\cite{stronginGergelBarkalovParGO}, \cite{strSergGO} --- \textit{the
-parallel computations are provided by means of simultaneous computing the values of the
+approach for solving the global optimization problems has been proposed \cite{stronginGergelBarkalovParGO}, \cite{strSergGO} --- \textit{the parallel computations are provided by means of simultaneous computing the values of the
 minimized function \(\varphi(y)\) at several different points within the search domain \(D\)}.
@@
 for the parallel computations on the high-performance computational systems with distributed memory is given.
-\subsection{Core Multidimensional Generalized Algorithm of Global Search}
+
+\subsection{Core Multidimensional Algorithm of Global Search}
 \label{subsec:corepar}
@@
 served as a basis for the development of a large number of efficient multiextremal optimization
-methods \cite{barkalovGergel2014}, \cite{Pizzuti}, \cite{gergel1996}, \cite{gergel1997}, \cite{grishaginStrongin1984}, \cite{sergeyev1995}, \cite{sergeyev1999}, \cite{sergeyevGrishagin2001}, \cite{sergeyevStronginLera2013}, \cite{Famularo2001}.
+methods \cite{barkalovGergel2014}, \cite{gergel1996}, \cite{gergel1997}, \cite{grishaginStrongin1984}, \cite{Pizzuti}, \cite{sergeyev1995}, \cite{sergeyev1999}, \cite{sergeyevGrishagin2001}, \cite{sergeyevStronginLera2013}, \cite{Famularo2001}.
 \par
@@
 \par
-The initial iteration of the algorithm is performed at an arbitrary point \(x^1(0,1)\).
+The initial iteration of the algorithm is performed at an arbitrary point \mbox{\(x^1\in(0,1)\)}.
 Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
-The selection of the trial point \(k+1\) for the next iteration is performed according to the following rules.
+The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the following rules.
 \par
@@
   \label{step1}
-0=x_0<x_1<...<x_{k+1}=1.
+0=x_0<x_1<...<x_{k+1}=1,
 \end{equation}
@@
 \begin{equation} \label{step2}
-m=\max_{1\leq i\leq k}\frac{|z_i-z_{i-1}|}{\rho_i}, \;
-M = \left\{
+M=\max_{1\leq i\leq k}\frac{|z_i-z_{i-1}|}{\rho_i}, \;
+m = \left\{
    \begin{array}{lr}
-     rm, & m > 0,\\
-     1, & m = 0,
+     rM, & M > 0,\\
+     1, & M = 0,
    \end{array}
@@
 \[
-R(i)=2\rho_i-4\frac{z_i}{M},\quad i=1,
+R(i)=2\rho_i-4\frac{z_i}{m},\quad i=1,
 \]
 \begin{equation} \label{step3}
-R(i)=\rho_i+\frac{(z_i-z_{i-1})^2}{M^2\rho_i}-2\frac{z_i+z_{i-1}}{M},\quad 1<i<k+1, \\
+R(i)=\rho_i+\frac{(z_i-z_{i-1})^2}{m^2\rho_i}-2\frac{z_i+z_{i-1}}{m},\quad 1<i<k+1, \\
 \end{equation}
 \[
-R(i)=2\rho_{i}-4\frac{z_{i-1}}{M},\quad i=k+1.
+R(i)=2\rho_{i}-4\frac{z_{i-1}}{m},\quad i=k+1.
 \]
@@
 \begin{equation} \label{step5}
-  x_{k+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{|z_{t}-z_{t-1}|}{M}\right]^N,\; \textrm{ if } 1<t<k+1,
+  x^{k+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{|z_{t}-z_{t-1}|}{M}\right]^N,\; \textrm{ if } 1<t<k+1,
 \end{equation}
 \[
-  x_{k+1}=\frac{x_t+x_{t-1}}{2},\; \textrm{ if } t=1 \textrm{ or } t=k+1.
+  x^{k+1}=\frac{x_t+x_{t-1}}{2},\; \textrm{ if } t=1 \textrm{ or } t=k+1.
 \]
@@
 \par
-The stop condition, which terminated the trials, is defined by the inequality
+The stopping condition, which terminated the trials, is defined by the inequality
 \begin{equation}
@@
 for the interval with the maximum characteristic from (\ref{step4}) and \(\varepsilon >0\) is the predefined
-accuracy of the optimization problem solution. If the stop condition is not satisfied,
-the index \(k\) is incremented by 1, and new global optimization iteration is executed.
+accuracy of the optimization problem solution. If the stopping condition is not satisfied,
+the index \(k\) is incremented by 1, and the new global optimization iteration is executed.
 \par
@@
 \par
-The convergence conditions for the described algorithm are given, for example, in \cite{strSergGO}.
+The convergence conditions of the described algorithm are given, for example, in \cite{strSergGO}.
 
@@
 Modern supercomputer computational systems consist of plenty of computational nodes,
-which include several multicore processors. In this case, random access memory is
-shared among the computational nodes --- the content of any memory location can be read
+which include several multicore processors. In this case, random access memory
+in the computational nodes is shared --- the content of any memory location can be read
 (written) by any computational core at any arbitrary moment. In most cases, shared memory is
@@
 \par
-\textit{Rule 4'}. Arrange the characteristics of the intervals computed according to (\ref{step3}) in the decreasing order
+\textit{Rule 4\'}. Arrange the characteristics of the intervals computed according to (\ref{step3}) in the decreasing order
 \begin{equation}
@@
 \par
-\textit{Rule 5'}. Execute new trials at the points \(x_{k+j},1\leq j\leq p\) located
+\textit{Rule 5\'}. Execute new trials at the points \(x^{k+j},1\leq j\leq p\) located
 in the intervals with the maximum characteristics from (\ref{step4par})
@@
 \begin{equation} \label{step5par}
-x_{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\; \textrm{ if } 1<t_j<k+1,
+x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\; \textrm{ if } 1<t_j<k+1,
 \end{equation}
 \[
-x_{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\; \textrm{ if } t_j=1 \textrm{ or } t_j=k+1.
+x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\; \textrm{ if } t_j=1 \textrm{ or } t_j=k+1.
 \]
 \par
-The stop condition (\ref{eq:stop_1}) should be checked for all intervals from (\ref{step4par}),
+The stopping condition (\ref{eq:stop_1}) should be checked for all intervals from (\ref{step4par}),
 where the scheduled trials are executed
@@
 \end{equation}
-If the stop condition is satisfied at least for one interval \(t_j, 1\le j\le p\), the calculations
+If the stopping condition is satisfied at least for one interval \(t_j, 1\le j\le p\), the calculations
 are terminated. Otherwise, the index \(k\) is incremented by \(p\), and new global optimization iteration is executed.
@@
 The set of mappings \(Y_s(x)\) from (\ref{eq:maps}) generates \(s\) information-linked one-dimensional
-problems of the kind (\ref{eq:oneDimTask}):
+problems (\ref{eq:oneDimTask}):
 \begin{equation}
@@
       \begin{displaymath}
-        x_i':y'=y^i(x_i'),1\le i\le s,
+        x_i':y'=y^i(x'),1\le i\le s,
       \end{displaymath}
@@
 \par
-The possibility of asynchronous data transfer (the computational nodes process the received
-data upon receipt only) is an important feature of the proposed parallel computation scheme.
+The possibility of asynchronous data transfer (the computational nodes process the obtained
+data after receiving only) is an important feature of the proposed parallel computation scheme.
 In addition, there is no any selected managing node within this scheme. The number of
@@
   \item Block 7 provides tuning the parallel computation schemes in the Globalizer system subject to the employed computational system architecture (the number of the computational cores, the availability of shared and/or distributed memory, the availability of computational accelerators, etc.) and the applied global optimization methods.
-  \item Block 8 is responsible for managing the parallel processes when performing the global optimization (determining the optimal configuration of parallel processes, distributing the processes between the computational unit, balancing the computational loads, etc.).
+  \item Block 8 is responsible for managing the parallel processes when performing the global optimization (determining the optimal configuration of parallel processes, distributing the processes between the computational units, balancing the computational loads, etc.).
   \item Block 9 is a managing subsystem, which controls the parallel computations when solving the global optimization problems.
@@
   \item Globalizer Pro is a research version for making various large-scale numerical experiments on the high-performance computational systems to investigate the new parallel global optimization methods being developed.
-  \item Globalizer Lite is a limited version for solving the global optimization problems with a medium complexity level, and is accessible free of charge.
+  \item Globalizer Lite is a limited version for solving the global optimization problems with a medium complexity level, and is accessible for free usage.
 \end{itemize}
@@
 problems with \textit{a priori} known properties: the number of local minima, the sizes of the attractors of these ones, the global minimum point, the function value at this point, etc.
-In order to simulate the computational costs inherent to the applied optimization problems, in all experiments computing the objective function was intentionally complicated by the auxiliary computations not altering the form of the function and the position of its minima.
-\par
-The numerical experiments were carried out first using the ExaMin system and then were reproduced using the Globalizer system.
+In order to simulate the computational costs that are inherent in the applied optimization problems, in all experiments computing the objective function was intentionally complicated by the auxiliary computations not altering the values of the function and the position of its minima.
+\par
 In all the executed experiments, 100 global optimization problems generated randomly by the GKLS-generator were solved.
@@
 \subsection{Comparison with over methods}
-The results of numerical comparison of three sequential algorithms --- DIRECT
+The results of numerical comparison of three serial algorithms --- DIRECT
 \cite{Jones}, DIRECT\textit{l} \cite{Gablonsky}, and MAGS from Section \ref{subsec:corepar} --- are presented below.
@@
 \begin{displaymath}
-    |y^k(i)-y^*(i)|\le \sqrt[N]{\Delta}(b(i)-a(i)), i\le i\le N
+    |y^k(i)-y^*(i)|\le \sqrt[N]{\Delta}(b(i)-a(i)), i\le i\le N,
 \end{displaymath}
-where \(N\) is the problem dimensionality, a and b are the borders of the search domain \(D\)
+where \(N\) is the problem dimensionality, $a$ and $b$ are the borders of the search domain \(D\)
 \footnote{In the case of MAGS, the stop condition was different slightly --- the
@@
 iterations $K_{max}$ was achieved. In this case, the value $K_{max}= 10^6$  was used to
-calculate the averaged number of iterations kav that corresponds to the lower estimate
+calculate the averaged number of iterations $k_{av}$ that corresponds to the lower estimate
 of this averaged value. The number of the unsolved problems is specified in the brackets.
@@
     \hline\noalign{\smallskip}
-     $N$ & Problem class & DIRECT & DIRECT\textit{l} & GSA \\
+     $N$ & Problem class & DIRECT & DIRECT\textit{l} & MAGS \\
     \noalign{\smallskip} \hline \noalign{\smallskip}
@@
 
-The results of the numerical experiments with GPMAGS with Intel Xeon Phi processors are presented in Tables \ref{tab:exResults} and \ref{tab:exResults2}.
+The results of the numerical experiments executed by using GPMAGS with Intel Xeon Phi processors are presented in Tables \ref{tab:4} and \ref{tab:5}.
 
@@
 The final series of experiments was executed using Xeon Phi processors. The results
-of these experiments are shown in row III. In this case, the speedup factors were
+of these experiments are shown in row III. In this case, the speedup is
 calculated relative to the results of the PMAGS method executed on the CPU using eight cores \((p=8)\).
@@
   \centering
-  \label{tab:exResults}
   \caption{Averaged numbers of iterations executed by GPMAGS for solving the test optimization problems}
+  \label{tab:4}
   \begin{tabular}{cccccccc}
@@
     \begin{center}
-    \textbf{Seriquential trial computations}
+    \textbf{Serial trial computations}
     \end{center}		}
@@
   \centering
-  \label{tab:exResults2}
   \caption{Speedup of parallel computations executed by GMAGS}
+  \label{tab:5}
   \begin{tabular}{cccccccc}
@@
     \begin{center}
-    \textbf{Serquential trial computations.}\\ \textit{Average number} \\ \textit{of iterations.}
+    \textbf{Serial trial computations.}\\ \textit{Average number} \\ \textit{of iterations}
     \end{center}		}
@@
 problems with the dimensions of \(N = 8\) and \(N = 10\): total 12 nodes with 36 graphic
-accelerators (three accelerators per the computational node) were employed (total 96,768 CUDA-cores were used).
+accelerators (three accelerators per the computational node) were employed (total 96768 CUDA-cores were used).
 \par
 The averaged execution time using the GPUs to solve the 8-dimensional problem was
-405,6 sec that was 5.9 times less than the execution time for CPU. The averaged time
-of solving the 10-dimensional problem was 2,055.8 sec. (in this case, the execution time for CPU
+405.6 sec that was 5.9 times less than the execution time for CPU. The averaged time
+of solving the 10-dimensional problem was 2055.8 sec (in this case, the execution time for CPU
 was not calculated because of the high complexity level of the required computations).
@@
 \bibitem{Barkalov2013}[10.1007/978-3-642-39958-9_14]
-\newblock K. Barkalov, A. Polov inkin, I. Meyerov, S. Sidorov, N. Zolotykh,
+\newblock K. Barkalov, A. Polovinkin, I. Meyerov, S. Sidorov, N. Zolotykh,
 \newblock SVM regression parameters optimization using parallel global search algorithm,
@@
 \bibitem{ciegisHentyParGO}[10.1007/978-0-387-09707-7]
-\newblock R. Ciegis, D. Henty, B. Kågström and J. Zilinskas,
+\newblock R. \v Ciegis, D. Henty, B. K\aa gstr\"om and J. \v Zilinskas,
 \newblock \emph{Parallel scientific computing and optimization: advances and applications},
@@
 \bibitem{Famularo1999}[10.1016/S0005-1098(99)00058-8]
-\newblock D. Famularo and P. Pugliese and Ya. D. Sergeyev,
+\newblock D. Famularo, P. Pugliese and Y. D. Sergeyev,
 \newblock A global optimization technique for checking parametric robustness,
@@
 \bibitem{fasanoPinter2013}[10.1007/978-1-4614-4469-5]
-\newblock G. Fasano and J. D. Pinter,
+\newblock G. Fasano and J. D. Pint\'er,
 \newblock \emph{Modeling and optimization in space engineering},
@@
 \bibitem{gavianoKvasovLeraSergeev2003}[10.1145/962437.962444]
-\newblock M. Gaviano and D.E. Kvasov and D. Lera and Ya.D. Sergeev,
+\newblock M. Gaviano, D. E. Kvasov, D. Lera and Y. D. Sergeev,
 \newblock Software for generation of classes of test functions with known local and global minima for global optimization,
@@
 \newblock V. P. Gergel,
-\newblock A global optimization algorithm for multivariate functions with lipschitzian first derivatives,
+\newblock A global optimization algorithm for multivariate functions with Lipschitzian first derivatives,
 \newblock \emph{J. Glob. Optim.}, \textbf{10(3)} (1997), 257--281.
@@
 \bibitem{Kvasov2015}[10.1016/j.advengsoft.2014.09.014]
-\newblock D. E. Kvasov and Ya. D. Sergeyev,
+\newblock D. E. Kvasov and Y. D. Sergeyev,
 \newblock Deterministic approaches for solving practical black-box global optimization problems,
@@
 \bibitem{Menniti}[10.1016/j.epsr.2007.10.009]
-\newblock D. E. Kvasov, D. Menniti, A. Pinnarelli, Ya. D. Sergeyev and N. Sorrentino,
+\newblock D. E. Kvasov, D. Menniti, A. Pinnarelli, Y. D. Sergeyev and N. Sorrentino,
 \newblock Tuning fuzzy power-system stabilizers in multi-machine systems by global optimization algorithms based on efficient domain partitions,
@@
 \bibitem{Pizzuti}[10.1007/s00211-002-0419-8]
-\newblock D. E. Kvasov, C. Pizzuti and Ya. D. Sergeyev,
+\newblock D. E. Kvasov, C. Pizzuti and Y. D. Sergeyev,
 \newblock Local tuning and partition strategies for diagonal go methods,
@@
 \bibitem{pardalosZhigljavskyZilinskas2016}[10.1007/978-3-319-29975-4]
-\newblock M. P. Pardalos, A. A. Zhigljavsky and J. Zilinskas,
+\newblock M. P. Pardalos, A. A. Zhigljavsky and J. \v Zilinskas,
 \newblock \emph{Advances in stochastic and deterministic global optimization},
@@
 \bibitem{pinterGO}[10.1007/978-1-4757-2502-5]
-\newblock J. D. Pinter,
+\newblock J. D. Pint\'er,
 \newblock \emph{Global optimization in action (Continuous and Lipschitz optimization: algorithms, implementations and applications)},
@@
 \bibitem{pinter2009}
-\newblock J. D. Pinter,
+\newblock J. D. Pint\'er,
 \newblock Software development for global optimization,
@@
 \bibitem{sergeyevKvasov2006}[	arXiv:1103.2056]
-\newblock 44. Ya.D. Sergeyev, D.E. Kvasov,
+\newblock Y. D. Sergeyev, D. E. Kvasov,
 \newblock Global search based on efficient diagonal partitions and a set of Lipschitz constants,
@@
 \bibitem{Famularo2001}[10.1023/A:1012391611462]
-\newblock Ya. D. Sergeyev, D. Famularo and P. Pugliese,
-\newblock Index branch-and-bound algorithm for lipschitz univariate global optimization with multiextremal constraints,
+\newblock Y. D. Sergeyev, D. Famularo and P. Pugliese,
+\newblock Index branch-and-bound algorithm for Lipschitz univariate global optimization with multiextremal constraints,
 \newblock \emph{J. Glob. Optim.}, \textbf{21(3)} (2001), 317--341.
@@
 \bibitem{strSergGO}[10.1007/978-1-4615-4677-1]
-\newblock R.G. Strongin and Ya.D. Sergeyev,
+\newblock R.G. Strongin and Y.D. Sergeyev,
 \newblock \emph{Global optimization with non-convex constraints. Sequential and parallel algorithms},
@@
 \bibitem{zilinskTornGO}[10.1007/3-540-50871-6]
-\newblock A. Törn and A. Zilinskas,
+\newblock A. T\"orn and A. \v Zilinskas,
 \newblock \emph{Global optimization},

### 2017-06-09 2017/NACO/paper.tex

@@
  % please put the address of the second  and third author
- \centerline{Lobachevsky State University of Nizhny Novgorod}
+ \centerline{Lobachevsky State University of Nizhni Novgorod}
    \centerline{23, Gagarin Avenue, Nizhni Novgorod, Russia}
@@
     \includegraphics[width=0.85\textwidth]{pictures/peanoC.eps}
-    \caption{A Peano curve approximation for the third density}
+    \caption{A Peano curve approximation for the third density level}
     \label{fig:peanoC}
@@
 \begin{equation} \label{step5}
-  x^{k+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{|z_{t}-z_{t-1}|}{M}\right]^N,\; \textrm{ if } 1<t<k+1,
+  x^{k+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{r|z_{t}-z_{t-1}|}{m}\right]^N,\; \textrm{ if } 1<t<k+1,
 \end{equation}
@@
 \begin{equation} \label{step5par}
-x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\; \textrm{ if } 1<t_j<k+1,
+x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{r|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\; \textrm{ if } 1<t_j<k+1,
 \end{equation}
@@
 \par
-The set of mappings \(Y_s(x)\) from (\ref{eq:maps}) generates \(s\) information-linked one-dimensional
-problems (\ref{eq:oneDimTask}):
+The set of mappings \(Y_s(x)\) from (\ref{eq:maps}) generates \(s\) information-linked one-\linebreak
+dimensional problems (\ref{eq:oneDimTask}):
 \begin{equation}
@@
 It is important to note that the family of the one-dimensional problems
-\(\varphi(y^l(x)),1 \leq l \leq s\) obtained as a result of the dimensionality
+$\varphi(y^l(x))$, $1 \leq l \leq s$ obtained as a result of the dimensionality
 reduction is an information-linked one --- the function values computed for any
@@
 problems with \textit{a priori} known properties: the number of local minima, the sizes of the attractors of these ones, the global minimum point, the function value at this point, etc.
-In order to simulate the computational costs that are inherent in the applied optimization problems, in all experiments computing the objective function was intentionally complicated by the auxiliary computations not altering the values of the function and the position of its minima.
+In order to simulate the computational costs that are inherent in the applied optimization problems, in all experiments computing the objective function values was intentionally complicated by the auxiliary computations not altering the values of the function and the position of its minima.
 \par

### 2017-06-23 2017/NACO/paper.tex

@@
 the optimization criterion can have several local optima within the search domain. 
-The existence of several local optima essentially makes
+The existence of several local optima makes
 finding the global optimum difficult essentially, since it requires examining the whole feasible
@@
 Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
-function \(\varphi(y(x))\), which satisfy the uniform H{\"o}lder condition i. e.
+function \(\varphi(y(x))\), which satisfy the uniform H{\"o}lder condition, i. e.
 \begin{equation}
@@
 \end{equation}
-the points \(x_0\), \(x_k\) were introduced additionally for the convenience of further
-explanation, the values of the minimized function \(z_0\), \(z_k\) at these points are undefined.
-\par
-\textit{Rule 2}. Compute current estimate of the H{\"o}lder constant \(H\) from (\ref{eq:holder})
+the points \(x_0\), \(x_{k+1}\) were introduced additionally for the convenience of further
+explanation, the values of the minimized function \(z_0\), \(z_{k+1}\) at these points are undefined.
+\par
+\textit{Rule 2}. Compute the current estimate of the H{\"o}lder constant \(H\) from (\ref{eq:holder})
 \begin{equation} \label{step2}
-M=\max_{1\leq i\leq k}\frac{|z_i-z_{i-1}|}{\rho_i}, \;
+M=\max_{1 < i\leq k}\frac{|z_i-z_{i-1}|}{\rho_i}, \;
 m = \left\{
@@
 expected location of the global minimum point. Thus, the rules (\ref{step4}) and (\ref{step5}) for selecting
-the interval for executing the next trial become more clear --- the point of every next
+the interval of the next trial become more clear --- the point of every next
 global optimization iteration is selected within the interval, where the global minimum
@@
 \begin{equation} \label{step5par}
-x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{r|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\; \textrm{ if } 1<t_j<k+1,
+x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{r|z_{t_j}-z_{t_j-1}|}{m}\right]^N,\; \textrm{ if } 1<t_j<k+1,
 \end{equation}
@@
     \end{itemize}
-    \item After completing any trial for any problem \(\varphi(y_l(x)),1\leq l\leq s\),
+    \item After completing any trial for any problem \(\varphi(y^l(x)),1\leq l\leq s\),
     at any point \(x'\in[0,1]\), it is necessary:
@@
     should check the queue of the received data; if there are any data in the queue,
-    these should be included into the search information. The MAGS algorithm uses these
-    accumulated data for the adaptive computing of the optimization iteration points (see Section \ref{subsec:corepar}).
+    these should be included into the search information. The PMAGS algorithm uses these
+    accumulated data for the adaptive computing of the optimization iteration points (see Section \ref{subsec:sharedpar}).
 \end{enumerate}
@@
 supercomputer management system – SLURM). Each supercomputer node had 2 Intel Sandy Bridge
-E5-2660 processors, 2.2 GHz, with 64 GB RAM. Each processor had 8 cores (i. e. total
+E5-2660 2.2 GHz processors with 64 GB RAM. Each processor had 8 cores (i. e. total
 16 CPU cores were available at each node). In order to generate the executable program code, the Intel C++ 14.0.2 compiler was used.
@@
 
-\subsection{Comparison with over methods}
+\subsection{Comparison with other methods}
 The results of numerical comparison of three serial algorithms --- DIRECT
@@
 \begin{displaymath}
-    |y^k(i)-y^*(i)|\le \sqrt[N]{\Delta}(b(i)-a(i)), i\le i\le N,
+    |y^k(i)-y^*(i)|\le \sqrt[N]{\Delta}(b(i)-a(i)), 1 \le i\le N,
 \end{displaymath}
@@
 
-In the second series of experiments, the parallel computations were executed using the CPUs.
+In the second series of experiments, the parallel computations were executed using PMAGS on CPUs.
 The achieved relative speedup in iterations is shown in row II; the speedup of the
-parallel computations is shown relative to the serial computations \((p=1)\).
+parallel computations is shown relatively to the serial computations \((p=1)\).
 
@@
 of these experiments are shown in row III. In this case, the speedup is
-calculated relative to the results of the PMAGS method executed on the CPU using eight cores \((p=8)\).
+calculated relatively to the results of the PMAGS method executed on the CPU using eight cores \((p=8)\).
 
@@
   \centering
-  \caption{Speedup of parallel computations executed by GMAGS}
+  \caption{Speedup of parallel computations executed by GPMAGS}
   \label{tab:5}
@@
 \bibitem{barkalovGergel2014}
-\newblock K. A. Barkalov and V. P. Gergel,
+\newblock K. Barkalov and V. Gergel,
 \newblock Multilevel scheme of dimensionality reduction for parallel global search algorithms,
@@
 \newblock K. Barkalov, V. Gergel and I. Lebedev,
-\newblock Use of XeonPhi coprocessor for solving global optimization problems,
+\newblock Use of Xeon Phi coprocessor for solving global optimization problems,
 \newblock \emph{LNCS}, \textbf{9251} (2015), 307--318.
@@
 \bibitem{gergel1993}
-\newblock V. P. Gergel,
+\newblock V. Gergel,
 \newblock A software system for multi-extremal optimization,
@@
 \bibitem{gergel1996}
-\newblock V. P. Gergel,
+\newblock V. Gergel,
 \newblock A method for using derivatives in the minimization of multiextremum functions,
@@
 \bibitem{gergel1997}[10.1023/A:1008290629896]
-\newblock V. P. Gergel,
+\newblock V. Gergel,
 \newblock A global optimization algorithm for multivariate functions with Lipschitzian first derivatives,
@@
 \bibitem{gergel2013}[doi:10.1016/j.procs.2013.05.164]
-\newblock V. P. Gergel, et al.,
+\newblock V. Gergel, et al.,
 \newblock High performance computing in biomedical applications,
@@
 \bibitem{gergel2015}[10.15866/ireaco.v8i1.4935]
-\newblock V. P. Gergel, et al.,
+\newblock V. Gergel, et al.,
 \newblock Recognition of surface defects of cold-rolling sheets based on method of localities,
@@
 \bibitem{gergelSidorov2015}[10.1007/978-3-319-21909-7_49]
-\newblock V. P. Gergel and S. V. Sidorov,
+\newblock V. Gergel and S. Sidorov,
 \newblock A two-level parallel global search algorithm for solving computationally intensive multi-extremal optimization problems,
@@
 \newblock D. E. Kvasov, C. Pizzuti and Y. D. Sergeyev,
-\newblock Local tuning and partition strategies for diagonal go methods,
+\newblock Local tuning and partition strategies for diagonal GO methods,
 \newblock \emph{Numerische Mathematik}, \textbf{94(1)} (2003), 93--106.
@@
 \newblock Multidimensional global optimization using the first derivatives,
-\newblock \emph{Comput. Math. Math. Phys.}, \textbf{39(5)} (1999), 743--752,.
+\newblock \emph{Comput. Math. Math. Phys.}, \textbf{39(5)} (1999), 743--752.
 
@@
 \bibitem{strSergGO}[10.1007/978-1-4615-4677-1]
-\newblock R.G. Strongin and Y.D. Sergeyev,
+\newblock R. G. Strongin and Y. D. Sergeyev,
 \newblock \emph{Global optimization with non-convex constraints. Sequential and parallel algorithms},
-\newblock Kluwer Academic Publishers, Dordrecht (2000).
+\newblock Kluwer Academic Publishers, Dordrecht (2000, 2nd ed. 2013, 3rd ed. 2014).
 

### 2017-06-26 2017/NACO/paper.tex

@@
 is highly important, since an estimate of the global minimum can be constructed on the
-base a finite number of computed values of the optimized function only in this case .
+basis of a finite number of computed values of the optimized function only in this case .
 \par
@@
 Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
-function \(\varphi(y(x))\), which satisfy the uniform H{\"o}lder condition, i. e.
+function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.
 \begin{equation}
@@
 are terminated. Otherwise, the index \(k\) is incremented by \(p\), and new global optimization iteration is executed.
-\par
+
+It should be pointed that in case of PMAGS $p$ trials are performed at each iteration (i.e., $k=np$, where $n$ is the number of executed iteration).
+
 The convergence conditions for the developed parallel algorithm are considered
@@
   \begin{enumerate}
-    \item Prior to beginning new trial at any point \(x'\in [0,1]\) for any problem \(\varphi(y^l(x)),1\leq l\leq s\), the following should be performed:
+    \item Prior to beginning each new trial at some point \(x'\in [0,1]\) for a problem \(\varphi(y^l(x)),1\leq l\leq s\), the following should be performed:
     \begin{itemize}
@@
     \end{itemize}
-    \item After completing any trial for any problem \(\varphi(y^l(x)),1\leq l\leq s\),
-    at any point \(x'\in[0,1]\), it is necessary:
+    \item After completing some trial for a problem \(\varphi(y^l(x)),1\leq l\leq s\),
+    at some point \(x'\in[0,1]\), it is necessary:
     \begin{itemize}
-      \item to determine all prototypes \(x_i',1\leq i\leq s\) for the point of the
+      \item to determine all prototypes \(x_i',1\leq i\leq s,\) for the point $x'\in[0,1]$ of the
       completed trial for each problem from the family (\ref{eq:oneDimProblemSerie}),
       \item to send the prototypes \(x_i',1\leq i\leq s\) and the result of the trial
-      \(z'=\varphi(y_l(x'))\) to all computational nodes to include the obtained data into
+      \(z'=\varphi(y^l(x'))\) to all computational nodes to include the obtained data into
       the search information processed according to the rules of the parallel global optimization algorithm.
@@
     \item Prior to starting the next global optimization iteration, the algorithm
-    should check the queue of the received data; if there are any data in the queue,
+    should check the queue of the received data; if there are some data in the queue,
     these should be included into the search information. The PMAGS algorithm uses these
@@
 \par
-A number of computational nodes had Intel Xeon Phi 5110P processors, each containing 60 cores
-(total 240 threads). In addition, the supercomputer included the computational nodes equipped
-with two NVIDIA Kepler K20X GPUs, each providing 14 stream multiprocessors (total 2688 CUDA-cores). In order to generate the executable code, the CUDA Toolkit 6.0 was used.
+A number of computational nodes had Intel Xeon Phi 5110P processors, each containing 60 cores. In addition, the supercomputer included the computational nodes equipped
+with two NVIDIA Kepler K20X GPUs, each providing 14 stream multiprocessors (in total 2688 CUDA-cores). In order to generate the executable code, the CUDA Toolkit 6.0 was used.
 \par
@@
 \begin{displaymath}
-    |y^k(i)-y^*(i)|\le \sqrt[N]{\Delta}(b(i)-a(i)), 1 \le i\le N,
+    |y^k_i-y^*_i|\le \sqrt[N]{\Delta}(b_i-a_i), 1 \le i\le N,
 \end{displaymath}
 where \(N\) is the problem dimensionality, $a$ and $b$ are the borders of the search domain \(D\)
-\footnote{In the case of MAGS, the stop condition was different slightly --- the
+\footnote{In the case of MAGS, the stopping condition was different slightly --- the
 global minimizer \(y^*\) was considered to be found, if MAGS generates a trial
@@
 problems with the dimensions of \(N = 8\) and \(N = 10\): total 12 nodes with 36 graphic
-accelerators (three accelerators per the computational node) were employed (total 96768 CUDA-cores were used).
+accelerators (three accelerators per the computational node) were employed (in total 96768 CUDA-cores were used).
 \par

### 2017-06-27 2017/NACO/paper.tex

@@
 It is important to note that the family of the one-dimensional problems
-$\varphi(y^l(x))$, $1 \leq l \leq s$ obtained as a result of the dimensionality
+$\varphi(y^l(x))$, $1 \leq l \leq s$, obtained as a result of the dimensionality
 reduction is an information-linked one --- the function values computed for any
@@
       \end{displaymath}
-      in accordance with the set of mappings \(y^l(x), 1\le l\le s\) for each problem from the family (\ref{eq:oneDimProblemSerie}),
-      \item send the prototypes \(x_i',1\leq i\leq s\) to all computational nodes in order
+      in accordance with the set of mappings \(y^l(x), 1\le l\le s\), for each problem from the family (\ref{eq:oneDimProblemSerie}),
+      \item send the prototypes \(x_i',1\leq i\leq s\), to all computational nodes in order
       to exclude the repeated selection of the intervals, which the prototypes fall in,
@@
       completed trial for each problem from the family (\ref{eq:oneDimProblemSerie}),
-      \item to send the prototypes \(x_i',1\leq i\leq s\) and the result of the trial
+      \item to send the prototypes \(x_i',1\leq i\leq s\), and the result of the trial
       \(z'=\varphi(y^l(x'))\) to all computational nodes to include the obtained data into

### 2017-07-01 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: update first 3 chapters

\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration

\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage[russian,english]{babel}
\DeclareMathOperator{\sign}{sign}
%
\begin{document}
%
\mainmatter              % start of the contributions
%
\title{Parallel Multi-objective Optimization on CPU Using Information Framework for Constructing Global Optimization Algorithms}
%
\titlerunning{Parallel Multi-objective Optimization}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Vladislav V. Sovrasov}
%
\authorrunning{Vladislav V. Sovrasov} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Vladislav V. Sovrasov}
%
\institute{State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
\email{sovrasov.vlad@gmail.com}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
В данной работе рассматривается параллельный алгоритм многокритериальной оптимизации. Рассматриваемый подход основан на применении информационно-статистического алгоритма к некоторой редуцированной однокритериальной задаче, множество глобальных оптимумов в которой совпадает с множеством слабоэффективных решений в исходной многокритериальной задаче. Последовательная версия данного метода была рассмотрена ранее. В данной работе к последовательному алгоритму многокритариальной оптимизации применяется схема распараллеливания по характеристикам, общая для всех информационно-статистических алгоритмов глобальной оптимизации. Также в работе впервые для многокритериального метода рассматривается одна из техник учёта локальных свойств оптимизируемой функции, позволяющая существенно ускорить сходимость.

\keywords{deterministi global optimization, multi-objective optimization, parallel numerical methods, derivative-free algorithms}
\end{abstract}
%
\section{Introduction}
\section{Problem Statement and Dimension Reduction}
Задача многокритериальной оптимизации ставится следующим образом:
\begin{equation}
  \label{eq:problem}
  \min\{f(y): y\in D\}, D=\{y\in \mathbb{R}^n: a_i \leqslant y_i \leqslant b_i, i=\overline{1,n} \}
\end{equation}
Будем считать, что компоненты вектор-функции (частные критерии) \(f_i(y), 1\leqslant i\leqslant m\), удовлетоворяют в \(D\) условию Липшица с константами \(L_i\):
\begin{displaymath}
\label{lip}
|f_i(y_1)-f_i(y_2)|\leqslant L_i\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L_i<\infty,i=\overline{1,m}
\end{displaymath}

As the solution to the problem (\ref{eq:problem}) usually accepted the set \(S(D)\in D\) of  strictly non-dominated points from the range of search, i. e.,
\begin{equation}
  \label{eq:slater}
  S(D) = \{y\in D: \nexists z\in D, f_i(z)<f_i(y),1\leqslant i \leqslant m\}
\end{equation}
which is usually referred as the set of semi-effective (or weakly effective) solutions. The conditions in the right-hand side of the definition (\ref{eq:slater}) are known as the principle of weak Pareto-optimality (or Slater's optimality principle).

The use of the evolvents \(y(x)\) i.e. the curves filling the space are a classic dimension-reduction scheme for global optimization algorithms \cite{evolvents2013}.
\begin{displaymath}
\label{cube}
\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
\end{displaymath}
\par
Such a mapping allows the reduction of a problem (\(\ref{eq:problem}\)) stated in a multidimensional space to solving a one-dimensional problem at the expense of worsening its properties.
In particular, the one-dimensional functions \(f_i(y(x))\) are not Lipschitzian but a Hölderian functions:
\begin{equation}
\label{eq:holder}
|f_i(y(x_1))-f_i(y(x_2))|\leqslant H_i{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
\end{equation}
where the Hölder constants \(H_i\) are related to the Lipschitz constant \(L_i\) by the relation
\begin{displaymath}
H_i=4L_id\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant n\}
\end{displaymath}
\par
Therefore, not limiting the generality, one can consider the solving of the
one-dimensional problem \(\min\{f(y(x)): x\in [0;1]\}\), satisfying Hölder condition. The issues of numerically building the mapping like a Peano curve and the corresponding theory have been considered in detail in \cite{evolvents2013}. Here we would note that an evolvent built numerically is an approximation to the theoretical Peano curve with a precision of the order \(2^{-m}\) where \(m\) is the building parameter of the evolvent.

\section{Description of the Parallel Algorithm With Local Refinement}
Рассмотрим схему скаляризации редуцированной задачи (\(\ref{eq:problem}\)), представленную в \cite{}. Пусть
\begin{equation}
  \varphi(x)=\max\{h(x,y):y\in [0;1]\},x\in [0;1].
\end{equation}
Рассмотрим скалярную задачу
\begin{equation}
  \label{eq:aux_problem}
  \varphi^*=\min\{\varphi(x):x\in [0;1]\}.
\end{equation}

Как показано в \cite{}, множество слабо-эффективных решений редуцированной задачи (\(\ref{eq:problem}\)) совпадает множеством глобально оптимальных решений задачи (\ref{eq:aux_problem}), т.е.
\begin{equation}
  \label{eq:s}
  S([0;1])=\{x\in [0;1]:\varphi(x)=\varphi^*\}
\end{equation}
Также в \cite{} показано, что функция \(\varphi(x)\) удовлетворяет условию Гёльдера при выполнении требований (\ref{equation}). Таким образом, к функции \(\varphi(x)\) можно применить информационно-статистический алгоритм глобального поиска, чтобы решить задачу (\ref{eq:aux_problem}). Однако, \(\varphi(x)\) задаётся через оператор \(\max\{...\}\), поэтому непосредственно вычислить её затруднительно. В \cite{} приведена модификация классического информационно-статистического алгоритма \cite{}, в которой значения \(\varphi(x)\) вычисляются приближённо. Далее приведём модифициованную версию указанного алгоритма. Модификация заключается в использовании техники local refinement, описанной в \cite{}, а также в распараллеливании по характеристикам \cite{}.

Первые две итерации производятся в концевых точках \(x^0=0\) и \(x^1=1\) интервала \([0;1]\). Выбор точек \(x^{k+j}, 1\leqslant j\leqslant p\) осуществляется по правилам:

Step 1. Renumber the points in the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which includes the boundary points of the interval \([0,1]\) as well as the points of preceding trials, by the lower indices in order of increasing coordinate values  i.e.
\begin{displaymath}
  0=x_0<x_1<\dotsc<x_{k+1}=1
\end{displaymath}

Step 2. Compute the values
\begin{equation}
\label{step2}
\mu_\nu=\max_{1\leqslant i\leqslant k}\dfrac{|f_\nu(x_i)-f_\nu(x_{i-1})|}{\Delta_i}, 1\leqslant \nu\leqslant m
\end{equation}

Step 3. Каждой точке \(x_i\), \(0\leqslant i\leqslant k\), сопоставить значение
\begin{equation}
  z_i=\max\{h(x_i,x_j):0\leqslant j\leqslant k\},
\end{equation}
где
\begin{equation}
  h(x_i,x_j)=\min\{\frac{f_\nu(x_i)-f_\nu(x_j)}{\mu_\nu}:1\leqslant \nu\leqslant m\}, 0\leqslant i,j\leqslant k
\end{equation}

Step 4. Для каждого интервала \((x_i,x_{i-1}),1\leqslant i\leqslant k\) вычислить величины
\begin{eqnarray}
  R(i) = \Delta_i + \frac{(z_i-z_{i-1})^2}{r^2\Delta_i}-\frac{z_i+z_{i-1}}{2r} \\
  R^*(i)=\frac{R(i)}{\sqrt{(z_i-z^*)(z_{i-1}-z^*)} + 1.5^{-\alpha}},
\end{eqnarray}
называемые характеристиками. При этом \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\), \(z^*=\min\{z_i:1\leqslant i\leqslant k\}\), а \(r>1\) и \(\alpha\in [10;30]\) --- параметры метода.

Step 5. Если \(q\not=0\) и \(s \mod q\not=0 \), то характеристики \(R(i)\), \(1 \leqslant i \leqslant k + 1\), упорядочить в порядке убывания
\begin{equation*}
  R(t_1) \geqslant R(t_2) \geqslant \dots \geqslant R(t_k) \geqslant R(t_{k+1})
\end{equation*}
и выбрать \(p\) наибольших характеристик с номерами интервалов \(t_j\), \(1 \leqslant j \leqslant p\). Иначе то же самое сделать с характеристиками \(R^*(i),1\leqslant i\leqslant k+1\). Здесь \(s\) --- номер текущей итерации. \(q\) --- параметр метода, отвечающий за степень интенсивности локального уточнения. Чем меньше \(q\), тем чаще используются характеристики \(R^*\), заставляющие метод выбирать следующие точки вблизи текущего найденного минимума.

Step 6. Провести новые испытания в точках \(x^{k+j}\), \(1 \leqslant j \leqslant p\):
\begin{equation}
  x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2} - \sign(z_{t_j} - z_{t_j-1})\frac{|z_{t_j} - z_{t_j-1}|^n}{2r}
\end{equation}
Все \(p\) испытаний на этом шаге могут быть произведены параллельно на \(p\) вычислительных устройствах.

The algorithm is terminated if the condition \(\Delta_{t_j}\leqslant \varepsilon\) is fulfilled at least for one of the numbers \(t_j\), \(1\leqslant j\leqslant p\); here \(\varepsilon >0\) is the predefined accuracy.
After the search is terminated, the set \(S(\{x^0,\dots ,x^k\})\) of all
non-dominated points of the truncated sequence \(\{x^0,\dots ,x^k\}\) is accepted as an estimation for \(S\) from (\ref{eq:s}).

The theoretical substantiation of this method when \(p=1\) and \(q=0\) is presented in \cite{}. Siffitient condition of convergence is: exists an iteration such that \(r\mu_\nu \geqslant 4H_\nu,1\leqslant \nu \leqslant m\).
\section{Experimental Results}
\section{Conclusion}
%
% ---- Bibliography ----
%
\begin{thebibliography}{5}
%
\bibitem {clar:eke}
Clarke, F., Ekeland, I.:
Nonlinear oscillations and
boundary-value problems for Hamiltonian systems.
Arch. Rat. Mech. Anal. 78, 315--333 (1982)

\bibitem {clar:eke:2}
Clarke, F., Ekeland, I.:
Solutions p\'{e}riodiques, du
p\'{e}riode donn\'{e}e, des \'{e}quations hamiltoniennes.
Note CRAS Paris 287, 1013--1015 (1978)

\bibitem {mich:tar}
Michalek, R., Tarantello, G.:
Subharmonic solutions with prescribed minimal
period for nonautonomous Hamiltonian systems.
J. Diff. Eq. 72, 28--55 (1988)

\bibitem {tar}
Tarantello, G.:
Subharmonic solutions for Hamiltonian
systems via a $\bbbz_{p}$ pseudoindex theory.
Annali di Matematica Pura (to appear)

\bibitem {rab}
Rabinowitz, P.:
On subharmonic solutions of a Hamiltonian system.
Comm. Pure Appl. Math. 33, 609--633 (1980)

\end{thebibliography}
\clearpage
\addtocmark[2]{Author Index} % additional numbered TOC entry
\renewcommand{\indexname}{Author Index}
\printindex
\clearpage
\end{document}

### 2017-07-02 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: overall updates, add references

@@
   \label{eq:problem}
-  \min\{f(y): y\in D\}, D=\{y\in \mathbb{R}^n: a_i \leqslant y_i \leqslant b_i, i=\overline{1,n} \}
+  \min\{f(y): y\in D\}, D=\{y\in \mathbb{R}^n: a_i \leqslant y_i \leqslant b_i, 1\leqslant i \leqslant n \}
 \end{equation}
@@
 \label{lip}
-|f_i(y_1)-f_i(y_2)|\leqslant L_i\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L_i<\infty,i=\overline{1,m}
+|f_i(y_1)-f_i(y_2)|\leqslant L_i\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L_i<\infty,1\leqslant i \leqslant m
 \end{displaymath}
@@
 \section{Description of the Parallel Algorithm With Local Refinement}
-Рассмотрим схему скаляризации редуцированной задачи (\(\ref{eq:problem}\)), представленную в \cite{}. Пусть
+\label{sec:algorithm}
+Рассмотрим схему скаляризации редуцированной задачи (\(\ref{eq:problem}\)), представленную в \cite{markinStrongin1993}. Пусть
 \begin{equation}
@@
 
-Как показано в \cite{}, множество слабо-эффективных решений редуцированной задачи (\(\ref{eq:problem}\)) совпадает множеством глобально оптимальных решений задачи (\ref{eq:aux_problem}), т.е.
+Как показано в \cite{markinStrongin1993}, множество слабо-эффективных решений редуцированной задачи (\(\ref{eq:problem}\)) совпадает множеством глобально оптимальных решений задачи (\ref{eq:aux_problem}), т.е.
 \begin{equation}
@@
 \end{equation}
-Также в \cite{} показано, что функция \(\varphi(x)\) удовлетворяет условию Гёльдера при выполнении требований (\ref{equation}).
+Также в \cite{markinStrongin1993} показано, что функция \(\varphi(x)\) удовлетворяет условию Гёльдера при выполнении требований (\ref{eq:holder}).
 Таким образом, к функции \(\varphi(x)\) можно применить информационно-статистический алгоритм глобального поиска, чтобы решить задачу (\ref{eq:aux_problem}).
 Однако, \(\varphi(x)\) задаётся через оператор \(\max\{...\}\), поэтому непосредственно вычислить её затруднительно.
-В \cite{} приведена модификация классического информационно-статистического алгоритма \cite{}, в которой значения \(\varphi(x)\) вычисляются приближённо.
+В \cite{markinStrongin1993} приведена модификация классического информационно-статистического алгоритма \cite{mixedAlg}, в которой значения \(\varphi(x)\) вычисляются приближённо.
 Далее приведём модифициованную версию указанного алгоритма.
-Модификация заключается в использовании техники local refinement, описанной в \cite{}, а также в распараллеливании по характеристикам \cite{}.
+Модификация заключается в использовании техники local refinement, описанной в \cite{strOptBook}(chapter 3), а также в распараллеливании по характеристикам \cite{strOptBook}(chapter 5).
 
@@
 
-The theoretical substantiation of this method when \(p=1\) and \(q=0\) is presented in \cite{}.
+The theoretical substantiation of this method when \(p=1\) and \(q=0\) is presented in \cite{strOptBook}(chapter 3).
 Siffitient condition of convergence is: exists an iteration such that \(r\mu_\nu \geqslant 4H_\nu,1\leqslant \nu \leqslant m\).
+\section{Test Problems}
+Для оценки степени ускорения сходимости модифицированного алгоритма из секции \ref{sec:algorithm} использовались следующие задачи:
+\begin{enumerate}
+  \item Markin-Strongin problem from \cite{markinStrongin1993}:
+    \begin{equation}
+      \left \{
+      \begin{array}{l}
+        f_1(y) = \min\{\sqrt{y_1^2+y_2^2},\sqrt{(y_1-1.5)^2+(y_2+1.5)^2}\} \\
+        f_2(y)=\sqrt{(y_1+0.5)^2+(y_2-0.5)^2} \\
+      \end{array}
+      \right .
+      y_1\in [-1;2],y_2\in [-2;1]
+    \end{equation}
+  \item Fonseca and Fleming problem from \cite{}:
+  \begin{equation}
+    \label{eq:fonseca}
+    \left \{
+    \begin{array}{l}
+      f_{1}\left(y\right) = 1 - \exp \left(-\sum_{i=1}^{n} \left(y_{i} - \frac{1}{\sqrt{n}} \right)^{2} \right) \\
+      f_{2}\left(y\right) = 1 - \exp \left(-\sum_{i=1}^{n} \left(y_{i} + \frac{1}{\sqrt{n}} \right)^{2} \right) \\
+    \end{array}
+    \right .
+    \quad y\in [-4;4]^n
+  \end{equation}
+  \item Schaffer function N. 2:
+    \begin{equation}
+      \left \{
+      \begin{array}{l}
+        f_1(y) = \begin{cases}
+                  -y,   & \text{if } y \leqslant 1 \\
+                   y-2, & \text{if } 1 < y \leqslant 3 \\
+                   4-y, & \text{if } 3 < y \leqslant 4 \\
+                   y-4, & \text{if } y > 4 \\
+                \end{cases} \\
+        f_2(y) = (y-5)^2\\
+      \end{array}
+      \right .
+      \quad y\in [-5;10]
+    \end{equation}
+  \item Poloni's two objective function:
+    \begin{equation}
+      \left \{
+      \begin{array}{l}
+        f_{1}\left(y\right) = \left[1 + \left(A_{1} - B_{1}\left(y\right) \right)^{2} + \left(A_{2} - B_{2}\left(y\right) \right)^{2} \right] \\
+        f_{2}\left(y\right) = \left(y_1 + 3\right)^{2} + \left(y_2 + 1 \right)^{2} \\
+      \end{array}
+      \right .
+      \quad y\in [-\pi;\pi]^2
+    \end{equation}
+    where
+    \begin{equation*}
+      \begin{cases}
+        A_{1} = 0.5 \sin \left(1\right) - 2 \cos \left(1\right) + \sin \left(2\right) - 1.5 \cos \left(2\right)  \\
+        A_{2} = 1.5 \sin \left(1\right) - \cos \left(1\right) + 2 \sin \left(2\right) - 0.5 \cos \left(2\right)  \\
+        B_{1}\left(y\right) = 0.5 \sin \left(y_1\right) - 2 \cos \left(y_1\right) + \sin \left(y_2\right) - 1.5 \cos \left(y_2\right)  \\
+        B_{2}\left(y\right) = 1.5 \sin \left(y_1\right) - \cos \left(y_1\right) + 2 \sin \left(y_2\right) - 0.5 \cos \left(y_2\right)
+      \end{cases}
+    \end{equation*}
+\end{enumerate}
 \section{Experimental Results}
+В данном разделе будем понимать под уcкорением работы метода ускорение по числу выполненных итераций, а не по времени выполнения.
+Если вычисление критериев задачи (\ref{eq:problem}) занимает достаточно много времени, то накладные затраты на реализацию решающих правил метода оптимизации невелики по сравнению с временем вычисления критериев \(f_i(y), 1\leqslant i\leqslant m\).
+
+\paragraph{Local refinement advantages.} В \cite{barkalovLebedev2016} приведены численные эксперименты, показывающие ускорение сходимости при применении техники локального уточнения в алгоритме, решающем скалярные задачи глобальной оптимизации.
+Подобных результатов стоит ожидать и от многокритериального алгоритма.
+Многокритериальный метод с локальным уточнением (MOLF) был применён к задаче Fonseca and Fleming (\ref{eq:fonseca}) при \(n=2\).
+Параметры метода были следующие: \(\varepsilon=0.01,\;r=4,\;q=4,\;\alpha=15,\;p=1\).
+До остановки метода было произведено 1176 итераций, количество найденных слабо оптимальных точек --- 90.
+При \(q=0\) (без локального уточнения) метод производит 1484 итерации, количество найденных слабо оптимальных точек --- 93.
+
+В дальнейшем для всех экспериментов будем использовать MOLF при \(q=4\).
+\paragraph{Parallel method results.}
 \section{Conclusion}
@@
 %
-\begin{thebibliography}{5}
-%
-\bibitem {clar:eke}
-Clarke, F., Ekeland, I.:
-Nonlinear oscillations and
-boundary-value problems for Hamiltonian systems.
-Arch. Rat. Mech. Anal. 78, 315--333 (1982)
+%\bibliographystyle{plain}
+%\bibliographystyle{alpha}
+\bibliographystyle{unsrt}
+%\bibliographystyle{splncs03}
+\bibliography{u_pdc_refs}
 
-\bibitem {clar:eke:2}
-Clarke, F., Ekeland, I.:
-Solutions p\'{e}riodiques, du
-p\'{e}riode donn\'{e}e, des \'{e}quations hamiltoniennes.
-Note CRAS Paris 287, 1013--1015 (1978)
-
-\bibitem {mich:tar}
-Michalek, R., Tarantello, G.:
-Subharmonic solutions with prescribed minimal
-period for nonautonomous Hamiltonian systems.
-J. Diff. Eq. 72, 28--55 (1988)
-
-\bibitem {tar}
-Tarantello, G.:
-Subharmonic solutions for Hamiltonian
-systems via a $\bbbz_{p}$ pseudoindex theory.
-Annali di Matematica Pura (to appear)
-
-\bibitem {rab}
-Rabinowitz, P.:
-On subharmonic solutions of a Hamiltonian system.
-Comm. Pure Appl. Math. 33, 609--633 (1980)
-
-\end{thebibliography}
 \clearpage

### 2017-07-04 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: add two pictures

@@
 \usepackage[russian,english]{babel}
+\usepackage{subfig}
+\usepackage{graphicx}
 \DeclareMathOperator{\sign}{sign}
@@
 \section{Test Problems}
+\label{sec:test_problems}
 Для оценки степени ускорения сходимости модифицированного алгоритма из секции \ref{sec:algorithm} использовались следующие задачи:
@@
 При \(q=0\) (без локального уточнения) метод производит 1484 итерации, количество найденных слабо оптимальных точек --- 93.
+На рис.
+\ref{fig:fonseca_slater} представлено множество \(S\) в задаче (\ref{eq:fonseca}) и его численные оценки, полученная при \(q=0\) (fig.
+\ref{fig:fonseca_slater}.a) и \(q=4\) (fig.
+\ref{fig:fonseca_slater}.b).
+
+\begin{figure}[ht]
+    \centering
+    \subfloat[MOLF with \(q=0\)]{{\includegraphics[width=.45\textwidth]{img/fonseca_glob.pdf} }}
+    \qquad
+    \subfloat[MOLF with \(q=4\)]{{\includegraphics[width=.45\textwidth]{img/fonseca_loc.pdf} }}
+    \caption{Numericas estimation of \(S\) obtained with MOLF}
+    \label{fig:fonseca_slater}
+\end{figure}
 
 В дальнейшем для всех экспериментов будем использовать MOLF при \(q=4\).
-\paragraph{Parallel method results.}
+\paragraph{Parallel method results.} Для каждой задачи из секции \ref{sec:test_problems}
+
 \section{Conclusion}

### 2017-07-06 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: add a reference to multi objective optimization test suite
u_pdc: update reference to the test problems source
u_pdc: add experiments results

@@
 
-\usepackage[T2A]{fontenc}
+\usepackage[T1,T2A]{fontenc}
 \usepackage[utf8]{inputenc}
+\usepackage[russian,english]{babel}
 \usepackage{amsmath,amssymb}
-\usepackage[russian,english]{babel}
 \usepackage{subfig}
 \usepackage{graphicx}
+\usepackage{tabularx}
 \DeclareMathOperator{\sign}{sign}
+\newcolumntype{K}[1]{>{\centering\arraybackslash}p{#1}}
 %
@@
     \end{equation}
-  \item Fonseca and Fleming problem from \cite{}:
+  \item Fonseca and Fleming problem \cite{Huband2006}:
   \begin{equation}
@@
   \end{equation}
-  \item Schaffer function N. 2:
+  \item Schaffer function N. 2 \cite{Huband2006}:
     \begin{equation}
+      \label{eq:schaffer2}
       \left \{
@@
     \end{equation}
-  \item Poloni's two objective function:
+  \item Poloni's two objective function \cite{Huband2006}:
     \begin{equation}
@@
     \end{equation*}
+%  \item A problem considered in \cite{}:
+%    \begin{equation}
+%      \left \{
+%      \begin{array}{l}
+%        f_1(y) = (y_1 - 1)y_2^2 +1 \\
+%        f_2(y)= y_2 \\
+%      \end{array}
+%      \right .
+%      \quad y\in [0;1]^2
+%    \end{equation}
 \end{enumerate}
 \section{Experimental Results}
+The computational experiments have been carried out on the Lobachevsky supercomputer at
+Lobachevsky State University of Nizhni Novgorod. A computational node included 2 Intel Sandy
+Bridge E5-2660 2.2 GHz processors, 64 Gb RAM. The CPUs had 8 cores (i. e. total 16 cores
+were available per a node).
+
 В данном разделе будем понимать под уcкорением работы метода ускорение по числу выполненных итераций, а не по времени выполнения.
@@
 На рис.
-\ref{fig:fonseca_slater} представлено множество \(S\) в задаче (\ref{eq:fonseca}) и его численные оценки, полученная при \(q=0\) (fig.
+\ref{fig:fonseca_slater} представлено множество \(S\) в задаче (\ref{eq:fonseca}) и его численные оценки, полученные при \(q=0\) (fig.
 \ref{fig:fonseca_slater}.a) и \(q=4\) (fig.
@@
     \centering
-    \subfloat[MOLF with \(q=0\)]{{\includegraphics[width=.45\textwidth]{img/fonseca_glob.pdf} }}
-    \qquad
-    \subfloat[MOLF with \(q=4\)]{{\includegraphics[width=.45\textwidth]{img/fonseca_loc.pdf} }}
-    \caption{Numericas estimation of \(S\) obtained with MOLF}
+    \subfloat[MOLF with \(q=0\)]{{\includegraphics[width=.5\textwidth]{img/fonseca_glob.pdf} }}
+    \subfloat[MOLF with \(q=4\)]{{\includegraphics[width=.5\textwidth]{img/fonseca_loc.pdf} }}
+    \caption{Numericas estimation of \(S\) obtained by MOLF}
     \label{fig:fonseca_slater}
@@
 В дальнейшем для всех экспериментов будем использовать MOLF при \(q=4\).
-\paragraph{Parallel method results.} Для каждой задачи из секции \ref{sec:test_problems}
+\paragraph{Parallel method results.} Чтобы продемонстрировать ускорение по итерациям, которое даёт MOLF при \(p > 1\), были решены все задачи из секции \ref{sec:test_problems} при значениях \(p=1,2,4,8,16\).
+Параметры метода при этом были следующие: \(r=4.5\), \(\varepsilon=0.01\), \(\alpha=15\).
+Для задачи (\ref{eq:schaffer2}) \(\varepsilon=0.001\).
+В таблице \ref{tab:exResultsIters} приведено количество итераций (в скобках мощность оценки множества \(S\)), а в таблице \ref{tab:exResultsItersSpeedup} --- ускорение по итерациям.
+
+Как видно из таблицы \ref{tab:exResultsIters}, мощность оценки множества \(S\) при изменении \(p\) меняется незначительно, т.
+е.
+качество полученной оценки сохраняется.
+В то же время, количество итераций соркащается пропорционально увеличению значения \(p\) (таблица \ref{tab:exResultsItersSpeedup}).
+На рис.
+\ref{fig:s_estimation_all} приведены примеры численных решений рассматриваемых задач.
+\begin{table}
+  \centering
+  \caption{Results of numerical experiments: number of iterations iterations}
+  \label{tab:exResultsIters}
+  \begin{tabular}{|l|K{1.7cm}|K{1.7cm}|K{1.7cm}|K{1.7cm}|K{1.7cm}|}
+\hline
+\textbf{Problem} & \multicolumn{5}{|c|}{\(p\)}\\
+\cline{2-6}
+  & \(1\) & \(2\) & \(4\) & \(8\) & \(16\)\\
+\hline
+Markin-Strongin & 1041(198) & 516(198) & 256(185) & 131(197) & 68(191) \\
+\hline
+Fonseca and Fleming 2d & 1181(93) & 636(99) & 386(111) & 176(95) & 106(97) \\
+\hline
+Fonseca and Fleming 3d & 5346(160) & 3551(183) & 1186(143) & 606(153) & 351(142) \\
+\hline
+Schaffer function N. 2 & 271(158) & 136(163) & 64(146) & 32(142) & 16(141)\\
+\hline
+Poloni's function & 3351(102) & 1706(90) & 856(88) & 426(96) & 201(99) \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}
+  \centering
+  \caption{Results of numerical experiments: speedup in iterations}
+  \label{tab:exResultsItersSpeedup}
+  \begin{tabular}{|l|K{1.5cm}|K{1.5cm}|K{1.5cm}|K{1.5cm}|}
+\hline
+\textbf{Problem} & \multicolumn{4}{|c|}{\(p\)}\\
+\cline{2-5}
+  & \(2\) & \(4\) & \(8\) & \(16\)\\
+\hline
+Markin-Strongin & 2.02 & 4.07 & 7.95 & 15.31 \\
+\hline
+Fonseca and Fleming 2d & 1.86 & 3.06 & 6.71 & 11.14 \\
+\hline
+Fonseca and Fleming 3d & 1.51 & 4.51 & 8.82 & 15.23 \\
+\hline
+Schaffer problem N. 2 & 1.99 & 4.23 & 8.47 & 16.94\\
+\hline
+Poloni's problem & 1.96 & 3.91 & 7.87 & 16.67 \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{figure}[ht]
+    \centering
+    \subfloat[Markin-Strongin problem]{{\includegraphics[width=.5\textwidth]{img/strongin.pdf} }}
+    \subfloat[Fonseca and Fleming 3d problem]{{\includegraphics[width=.5\textwidth]{img/fonseca.pdf} }}
+
+    \subfloat[Schaffer problem N. 2]{{\includegraphics[width=.5\textwidth]{img/schaffer2.pdf} }}
+    \subfloat[Poloni's problem]{{\includegraphics[width=.5\textwidth]{img/poloni.pdf} }}
+    \caption{Numericas estimation of \(S\) obtained by MOLF}
+    \label{fig:s_estimation_all}
+\end{figure}
+
+\paragraph{Parallel method time speedup.} Целевые функции всех перечисленных в разделе задачи обладают невысокой вычислительной сложностью.
+Чтобы продемонстрировать возможность получения ускорения по времени для задач с трудновычислимыми критериями, в критерии были внесены дополнительные интенсивные вычисления с плавающей точкой, не влияющие на результирующие значения.
+Полученные при этом ускорения приведены в таблице \ref{}.
 
 \section{Conclusion}
+
 %

### 2017-07-06 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: update conclusion section

@@
 Чтобы продемонстрировать возможность получения ускорения по времени для задач с трудновычислимыми критериями, в критерии были внесены дополнительные интенсивные вычисления с плавающей точкой, не влияющие на результирующие значения.
-Полученные при этом ускорения приведены в таблице \ref{}.
+Полученные при этом ускорения приведены в таблице \ref{tab:exResultsTimeSpeedup}.
+Ускорение по времени во всех случаях меньше, чем ускорение по итерациям, т.к.
+часть времени выполнения занимает работа метоад оптимизации, однако распареллеливание в большинстве случаев довольно эффективно и в случае 16 рабочих потоков.
+При ещё большей вычислительной сложности критериев стоит ожидать увеличения ускорения по времени для большого числа потоков.
+
+\begin{table}[ht]
+  \centering
+  \caption{Results of numerical experiments: speedup in time}
+  \label{tab:exResultsTimeSpeedup}
+  \begin{tabular}{|l|K{1.5cm}|K{1.5cm}|K{1.5cm}|K{1.5cm}|K{1.5cm}|}
+\hline
+\textbf{Problem} & \multicolumn{5}{|c|}{\(p\)}\\
+\cline{2-6}
+&\(1\)(time, s) & \(2\) & \(4\) & \(8\) & \(16\)\\
+\hline
+Markin-Strongin & 104.47 & 1.97 & 3.65 & 6.79 & 9.90 \\
+\hline
+Fonseca and Fleming 2d & 118.95 & 1.85 & 2.81 & 5.79 & 6.40 \\
+\hline
+Fonseca and Fleming 3d & 554.45 & 1.51 & 4.14 & 8.05 & 10.69 \\
+\hline
+Schaffer problem N. 2 & 27.21 & 1.91 & 3.79 & 6.39 & 6.37\\
+\hline
+Poloni's problem & 336.74 & 1.82 & 3.64 & 6.98 & 10.60 \\
+\hline
+\end{tabular}
+\end{table}
 
 \section{Conclusion}
-
+В работе был рассмотрен метод решения многокритериальных задач оптимизации, позволяющий находить равномерную оценку множества слабо-оптимальных точек.
+К нему применены техники распараллеливания и ускорения сходимости, общие для всех алгоритмов подобного структуры \cite{}.
+Проведённые численные эксперименты показали наличие ускорения сходимости при использовании local refinement.
+Также для данного метода оказалось эффективным распареллеливание по характеристикам.
+Как и для информационных методов скалярной оптимизации, ускорение по итерациям при распареллеливании по характеристикам оказывается линейным в большинстве случаев \cite{}.
+Исходя из результатов эскпериментов, параллельный MOLF применим в задачах низкой размерности с трудновычислимыми критериями.
 %

### 2017-07-07 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: fix Makefile, update refs

@@
 \hline
-\textbf{Problem} & \multicolumn{5}{|c|}{\(p\)}\\
+\textbf{Problem} & \multicolumn{5}{c|}{\(p\)}\\
 \cline{2-6}
@@
 \hline
-\textbf{Problem} & \multicolumn{4}{|c|}{\(p\)}\\
+\textbf{Problem} & \multicolumn{4}{c|}{\(p\)}\\
 \cline{2-5}
@@
 В работе был рассмотрен метод решения многокритериальных задач оптимизации, позволяющий находить равномерную оценку множества слабо-оптимальных точек.
-К нему применены техники распараллеливания и ускорения сходимости, общие для всех алгоритмов подобного структуры \cite{}.
+К нему применены техники распараллеливания и ускорения сходимости, общие для всех алгоритмов подобного структуры \cite{strOptBook}, chapter 3.
 Проведённые численные эксперименты показали наличие ускорения сходимости при использовании local refinement.
 Также для данного метода оказалось эффективным распареллеливание по характеристикам.
-Как и для информационных методов скалярной оптимизации, ускорение по итерациям при распареллеливании по характеристикам оказывается линейным в большинстве случаев \cite{}.
+Как и для информационных методов скалярной оптимизации, ускорение по итерациям при распареллеливании по характеристикам оказывается линейным в большинстве случаев \cite{barkalovLebedef2016}.
 Исходя из результатов эскпериментов, параллельный MOLF применим в задачах низкой размерности с трудновычислимыми критериями.

### 2017-07-07 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: add introduction, cosmetic changes in other sections

@@
 \section{Introduction}
+Задачи многокритериальной оптимизации привлекают в последнее время всё больший интерес, т.к.
+наиболее полно отражают противоречивость требований, возникающих к моделируемым оъбектам реального мира.
+В таких задачах ответом является одно или множество компромиссных решений.
+Как правило, эти компромиссные решения выбираются из множества, в котором все элементы не могут быть улучшены ни по одному критерию без ухудшения других (Pareto-optimal or efficient solutions) или из множества, в котором ни один элемент не может быть улучшен по всем криериям сразу (Slater-optimal or weak-efficient solutions).
+Последнее требование мягче и все Pareto-optimal solutions являются и Slater-optimal.
+
+Найти всё множество efficient или weak-efficient точек является наиболее вычислительно сложной задачей, однако это даёт широкий выбор возможных компромиссных решений.
+Разработано множество подходов к поиску этих множеств: применение генетических алгоритмов непосредственно к многокритериалной задаче \cite{}, использование параметрических свёрток, сводящих задачу к серии скалярных задач глобально оптимизации \cite{}, различные подходы для решения задач с выпуклыми или линейными критериями \cite{}.
+Основной проблемой при исользовании большинства указанных методов является отсутствие теоретической гарантии равномерной сходимости ко всему искомому множеству на широком классе задач.
+Метод, обладающий этим свойством в случае, когда все критерии удовлетворяют условию Липшица, был разработан на основе информационного метода глобальной оптимимзации \cite{}.
+В статье будет рассмотрена его параллельная и ускоренная с помощью техники local refinement \cite{strOptBook}(chapter 3) модификация.
+
 \section{Problem Statement and Dimension Reduction}
@@
 
-Step 1.
-Renumber the points in the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which includes the boundary points of the interval \([0,1]\) as well as the points of preceding trials, by the lower indices in order of increasing coordinate values  i.e.
+\textit{Step 1.} Renumber the points in the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which includes the boundary points of the interval \([0,1]\) as well as the points of preceding trials, by the lower indices in order of increasing coordinate values  i.e.
 \begin{displaymath}
@@
 
-Step 2. Compute the values
-\begin{equation}
-\label{step2}
-\mu_\nu=\max_{1\leqslant i\leqslant k}\dfrac{|f_\nu(x_i)-f_\nu(x_{i-1})|}{\Delta_i}, 1\leqslant \nu\leqslant m
-\end{equation}
+\textit{Step 2.} Compute the lover bound \(\mu_\nu\) for the Lipzhitz constant for each objective function \(f_\nu(x),1\leqslant\nu\leqslant m\):
+\begin{gather}
+\label{eq:step2_1}
+\mu_\nu=\max_{1\leqslant i\leqslant k}\dfrac{|f_\nu(x_i)-f_\nu(x_{i-1})|}{\Delta_i}, 1\leqslant \nu\leqslant m \\
+\label{eq:step2_2}
+\Delta_i=(x_i-x_{i-1})^\frac{1}{N}
+\end{gather}
 
-Step 3. Каждой точке \(x_i\), \(0\leqslant i\leqslant k\), сопоставить значение
+\textit{Step 3.} To each point \(x_i\), \(0\leqslant i\leqslant k\), juxtapose the value
 \begin{equation}
@@
 \end{equation}
-где
+where
 \begin{equation}
@@
 
-Step 4. Для каждого интервала \((x_i,x_{i-1}),1\leqslant i\leqslant k\) вычислить величины
+\textit{Step 4.} For each interval \((x_i,x_{i-1}),1\leqslant i\leqslant k\) compute two characteristics:
 \begin{eqnarray}
@@
 \end{eqnarray}
-называемые характеристиками. При этом \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\), \(z^*=\min\{z_i:1\leqslant i\leqslant k\}\), а \(r>1\) и \(\alpha\in [10;30]\) --- параметры метода.
+where \(\Delta_i\) is from (\ref{eq:step2_2}), \(z^*=\min\{z_i:1\leqslant i\leqslant k\}\). \(r>1\) and \(\alpha\in [10;30]\) are input parameters of the algorithm.
 
-Step 5. Если \(q\not=0\) и \(s \mod q\not=0 \), то характеристики \(R(i)\), \(1 \leqslant i \leqslant k + 1\), упорядочить в порядке убывания
+\textit{Step 5.} Если \(q\not=0\) и \(s \mod q\not=0 \), то характеристики \(R(i)\), \(1 \leqslant i \leqslant k + 1\), упорядочить в порядке убывания
 \begin{equation*}
@@
 
-Step 6. Провести новые испытания в точках \(x^{k+j}\), \(1 \leqslant j \leqslant p\):
+\textit{Step 6.} Провести новые испытания в точках \(x^{k+j}\), \(1 \leqslant j \leqslant p\):
 \begin{equation}
@@
 
-\paragraph{Local refinement advantages.} В \cite{barkalovLebedev2016} приведены численные эксперименты, показывающие ускорение сходимости при применении техники локального уточнения в алгоритме, решающем скалярные задачи глобальной оптимизации.
+\paragraph{Local refinement advantages.} В \cite{barkalovLebedev2016} приведены численные эксперименты, показывающие ускорение сходимости при применении техники local refinement в алгоритме, решающем скалярные задачи глобальной оптимизации.
 Подобных результатов стоит ожидать и от многокритериального алгоритма.
-Многокритериальный метод с локальным уточнением (MOLF) был применён к задаче Fonseca and Fleming (\ref{eq:fonseca}) при \(n=2\).
+Multi-objective algorithm with local refinement (MOALR) был применён к задаче Fonseca and Fleming (\ref{eq:fonseca}) при \(n=2\).
 Параметры метода были следующие: \(\varepsilon=0.01,\;r=4,\;q=4,\;\alpha=15,\;p=1\).
@@
 \ref{fig:fonseca_slater}.b).
+Из рис.
+\ref{fig:fonseca_slater} также видно, что приближённое решение покрывает всё множество weak-optimal solutions.
 
@@
     \centering
-    \subfloat[MOLF with \(q=0\)]{{\includegraphics[width=.5\textwidth]{img/fonseca_glob.pdf} }}
-    \subfloat[MOLF with \(q=4\)]{{\includegraphics[width=.5\textwidth]{img/fonseca_loc.pdf} }}
-    \caption{Numericas estimation of \(S\) obtained by MOLF}
+    \subfloat[MOALR with \(q=0\)]{{\includegraphics[width=.5\textwidth]{img/fonseca_glob.pdf} }}
+    \subfloat[MOALR with \(q=4\)]{{\includegraphics[width=.5\textwidth]{img/fonseca_loc.pdf} }}
+    \caption{Numericas estimation of \(S\) obtained by MOALR}
     \label{fig:fonseca_slater}
@@
 
-В дальнейшем для всех экспериментов будем использовать MOLF при \(q=4\).
-\paragraph{Parallel method results.} Чтобы продемонстрировать ускорение по итерациям, которое даёт MOLF при \(p > 1\), были решены все задачи из секции \ref{sec:test_problems} при значениях \(p=1,2,4,8,16\).
+В дальнейшем для всех экспериментов будем использовать MOALR при \(q=4\).
+\paragraph{Parallel method results.} Чтобы продемонстрировать ускорение по итерациям, которое даёт MOALR при \(p > 1\), были решены все задачи из секции \ref{sec:test_problems} при значениях \(p=1,2,4,8,16\).
 Параметры метода при этом были следующие: \(r=4.5\), \(\varepsilon=0.01\), \(\alpha=15\).
@@
     \subfloat[Poloni's problem]{{\includegraphics[width=.5\textwidth]{img/poloni.pdf} }}
-    \caption{Numericas estimation of \(S\) obtained by MOLF}
+    \caption{Numericas estimation of \(S\) obtained by MOALR}
     \label{fig:s_estimation_all}
@@
 В работе был рассмотрен метод решения многокритериальных задач оптимизации, позволяющий находить равномерную оценку множества слабо-оптимальных точек.
-К нему применены техники распараллеливания и ускорения сходимости, общие для всех алгоритмов подобного структуры \cite{strOptBook}, chapter 3.
+К нему применены техники распараллеливания и ускорения сходимости, общие для всех алгоритмов подобного структуры.
 Проведённые численные эксперименты показали наличие ускорения сходимости при использовании local refinement.
 Также для данного метода оказалось эффективным распареллеливание по характеристикам.
-Как и для информационных методов скалярной оптимизации, ускорение по итерациям при распареллеливании по характеристикам оказывается линейным в большинстве случаев \cite{barkalovLebedef2016}.
-Исходя из результатов эскпериментов, параллельный MOLF применим в задачах низкой размерности с трудновычислимыми критериями.
+Как и для информационных методов скалярной оптимизации, ускорение по итерациям при распареллеливании по характеристикам оказывается линейным в большинстве случаев (см.
+результаты распараллеливания метода скалярной оптимизации напр., в \cite{barkalovLebedef2016}).
+Исходя из результатов эскпериментов, параллельный MOALR целесообразно применять в задачах низкой размерности (1---3) с трудновычислимыми критериями.
+Кроме того, на численных примерах продемонстрировано свойство равномерной сходимости рассматриваемого алгоритма ко всему множеству weak-optimal solutions.
 %

### 2017-07-10 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: translated version

--- 2017-07-07 2017/URAL-PDC/u_pdc_paper.tex
@@
 
-\usepackage[T1,T2A]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[russian,english]{babel}
+\usepackage[english]{babel}
 \usepackage{amsmath,amssymb}
@@
 %
-\title{Parallel Multi-objective Optimization on CPU Using Information Framework for Constructing Global Optimization Algorithms}
+\title{Parallel Multi-objective Optimization on CPU Using Information Framework for
+Constructing Global Optimization Algorithms}
 %
@@
 \begin{abstract}
-В данной работе рассматривается параллельный алгоритм многокритериальной оптимизации.
-Рассматриваемый подход основан на применении информационно-статистического алгоритма к некоторой редуцированной однокритериальной задаче, множество глобальных оптимумов в которой совпадает с множеством слабоэффективных решений в исходной многокритериальной задаче.
-Последовательная версия данного метода была рассмотрена ранее.
-В данной работе к последовательному алгоритму многокритариальной оптимизации применяется схема распараллеливания по характеристикам, общая для всех информационно-статистических алгоритмов глобальной оптимизации.
-Также в работе впервые для многокритериального метода рассматривается одна из техник учёта локальных свойств оптимизируемой функции, позволяющая существенно ускорить сходимость.
+In the present work, a parallel algorithm for the multi-objective optimization is considered. The
+considered approach is based on the application of the information-statistical algorithm to some
+reduced unicriterial problem, the set of the global optima in which coincides to the set of the
+weakly efficient solutions of the initial multi-objective problem. The sequential version of this
+method has been considered earlier. In the present work, the parallelization scheme based on
+the characteristics, which is common for all information-statistical global search algorithms,
+was applied to the sequential algorithm of the multi-objective optimization. Also, in the present
+work one of the techniques of accounting for the local properties of the optimized function
+allowing accelerating the convergence essentially has been considered for the multi-objective
+method for the first time.
 
-\keywords{deterministi global optimization, multi-objective optimization, parallel numerical methods, derivative-free algorithms}
+\keywords{deterministic global optimization, multi-objective optimization, parallel numerical
+methods, derivative-free algorithms}
 \end{abstract}
@@
 \section{Introduction}
-\section{Problem Statement and Dimension Reduction}
-Задача многокритериальной оптимизации ставится следующим образом:
+The multi-objective optimization problems attract an increasing interest in recent years since
+these ones reflect the controversy of the requirements arising to the modeled objects of real
+world the most completely. In such problems, the solutions are a compromised decision of a set
+of these ones. As a rule, these compromised solutions are chosen either from a set, in which all
+elements cannot be improved with respect to any criterion without the worsening of the other
+criteria (Pareto-optimal or efficient solutions) or from a set, in which no one element could be
+improved with respect to all criteria at once (Slater-optimal or weakly efficient solutions). The
+latter requirement is weaker, and all Pareto-optimal solutions are the Slater-optimal ones as
+well.
+
+To find the complete set of efficient or weakly efficient points is the most computationally
+complex problem, however it provides a wide choice of possible compromised solutions. A
+plenty of approaches to the finding of these sets has been developed: the application of the
+genetic algorithms to the multi-objective problem directly~\cite{DebPratap2002}, the use of the
+parametric convolutions reducing the problem to a series of the scalar global optimization
+problems~\cite{GergelKozinov2017}, various approaches to the solving the problems with the
+convex~ \cite{ZhangZuo2013} or linear~\cite{Benson1998} criteria. The main problem in the
+using of majority of these methods is the absence of the theoretical guarantee of the uniform
+convergence to the whole sought set of the Pareto-optimal solutions on a wide class of
+problems including the ones with the non-convex multiextremal criteria. A method having this
+property in the case, when all criteria satisfy the Lipschitz condition has been developed on the
+basis of the information global search algorithm~\cite{markinStrongin1993}. In the present
+paper, the parallel and accelerated by the local refinement technique~\cite{strOptBook}(chapter
+3) modification of this method will be considered.
+
+\section{Problem Statement and Dimensionality Reduction}
+The multi-objective optimization problems are stated in the following way:
 \begin{equation}
   \label{eq:problem}
-  \min\{f(y): y\in D\}, D=\{y\in \mathbb{R}^n: a_i \leqslant y_i \leqslant b_i, 1\leqslant i \leqslant n \}
+  \min\{f(y): y\in D\}, D=\{y\in \mathbb{R}^n: a_i \leqslant y_i \leqslant b_i, 1\leqslant i
+\leqslant n \}
 \end{equation}
-Будем считать, что компоненты вектор-функции (частные критерии) \(f_i(y), 1\leqslant i\leqslant m\), удовлетоворяют в \(D\) условию Липшица с константами \(L_i\):
+Let us assume the components of the vector function (the partial criteria) \(f_i(y), 1\leqslant
+i\leqslant m\) to satisfy the Lipschitz condition with the Lipschitz constant \(L_i\) in \(D\):
 \begin{displaymath}
 \label{lip}
-|f_i(y_1)-f_i(y_2)|\leqslant L_i\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L_i<\infty,1\leqslant i \leqslant m
+|f_i(y_1)-f_i(y_2)|\leqslant L_i\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L_i<\infty,1\leqslant i
+\leqslant m
 \end{displaymath}
 
-As the solution to the problem (\ref{eq:problem}) usually accepted the set \(S(D)\in D\) of  strictly non-dominated points from the range of search, i. e.,
+The set \(S(D)\in D\) of strictly non-dominated points from the search domain is accepted as the
+solution of the problem (\ref{eq:problem}) usually i. e.
+
 \begin{equation}
@@
 \end{equation}
-which is usually referred as the set of semi-effective (or weakly effective) solutions.
-The conditions in the right-hand side of the definition (\ref{eq:slater}) are known as the principle of weak Pareto-optimality (or Slater's optimality principle).
 
-The use of the evolvents \(y(x)\) i.e. the curves filling the space are a classic dimension-reduction scheme for global optimization algorithms \cite{evolvents2013}.
+which is usually referred as the set of semi-efficient (or weakly efficient) solutions. The
+conditions in the right-hand side of the definition (\ref{eq:slater}) are known as the principle of
+weak Pareto-optimality (or Slater's optimality principle).
+
+The use of the evolvents \(y(x)\) i. e. the space-filling curves is a classic dimensionality
+reduction scheme for global optimization algorithms~\cite{evolvents2013}.
 \begin{displaymath}
 \label{cube}
-\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
+\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant
+N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
 \end{displaymath}
 \par
-Such a mapping allows the reduction of a problem (\(\ref{eq:problem}\)) stated in a multidimensional space to solving a one-dimensional problem at the expense of worsening its properties.
-In particular, the one-dimensional functions \(f_i(y(x))\) are not Lipschitzian but a Hölderian functions:
+Such a mapping allows the reduction of the problem (\(\ref{eq:problem}\)) stated in a
+multidimensional space to solving a one-dimensional problem at the expense of worsening its
+properties.
+In particular, the one-dimensional functions \(f_i(y(x))\) are not Lipschitzian but the
+H\"olderian functions:
 \begin{equation}
@@
 \end{equation}
-where the Hölder constants \(H_i\) are related to the Lipschitz constant \(L_i\) by the relation
+where the H\"older constants \(H_i\) are related to the Lipschitz constant \(L_i\) by the relation
 \begin{displaymath}
@@
 Therefore, not limiting the generality, one can consider the solving of the
-one-dimensional problem \(\min\{f(y(x)): x\in [0;1]\}\), satisfying Hölder condition.
-The issues of numerically building the mapping like a Peano curve and the corresponding theory have been considered in detail in \cite{evolvents2013}.
-Here we would note that an evolvent built numerically is an approximation to the theoretical Peano curve with a precision of the order \(2^{-m}\) where \(m\) is the building parameter of the evolvent.
+one-dimensional problem \(\min\{f(y(x)): x\in [0;1]\}\), satisfying the H\"older condition. The
+issues of the numerical building the mapping like a Peano curve and the corresponding theory
+have been considered in detail in~\cite{evolvents2013}. Here we would note that an evolvent
+built numerically is an approximation of the theoretical Peano curve with a precision of the
+order \(2^{-m}\) where \(m\) is the building parameter of the evolvent.
 
@@
 \label{sec:algorithm}
-Рассмотрим схему скаляризации редуцированной задачи (\(\ref{eq:problem}\)), представленную в \cite{markinStrongin1993}. Пусть
+Let us consider a scalarization scheme for the reduced problem (\(\ref{eq:problem}\)),
+presented in~\cite{markinStrongin1993}. Let
 \begin{equation}
@@
 \end{equation}
-Рассмотрим скалярную задачу
+Let us consider a scalar problem
 \begin{equation}
@@
 
-Как показано в \cite{markinStrongin1993}, множество слабо-эффективных решений редуцированной задачи (\(\ref{eq:problem}\)) совпадает множеством глобально оптимальных решений задачи (\ref{eq:aux_problem}), т.е.
+As it has been shown in~\cite{markinStrongin1993}, the set of weakly efficient solutions of the
+reduced problem (\(\ref{eq:problem}\)) coincides to the set of the globally optimal solutions of
+the problem (\ref{eq:aux_problem}) i. e.
 \begin{equation}
@@
 \end{equation}
-Также в \cite{markinStrongin1993} показано, что функция \(\varphi(x)\) удовлетворяет условию Гёльдера при выполнении требований (\ref{eq:holder}).
-Таким образом, к функции \(\varphi(x)\) можно применить информационно-статистический алгоритм глобального поиска, чтобы решить задачу (\ref{eq:aux_problem}).
-Однако, \(\varphi(x)\) задаётся через оператор \(\max\{...\}\), поэтому непосредственно вычислить её затруднительно.
-В \cite{markinStrongin1993} приведена модификация классического информационно-статистического алгоритма \cite{mixedAlg}, в которой значения \(\varphi(x)\) вычисляются приближённо.
-Далее приведём модифициованную версию указанного алгоритма.
-Модификация заключается в использовании техники local refinement, описанной в \cite{strOptBook}(chapter 3), а также в распараллеливании по характеристикам \cite{strOptBook}(chapter 5).
+Also, it has been shown in~\cite{markinStrongin1993} that the function \(\varphi(x)\) satisfies
+the H\"older condition when the requirements (\ref{eq:holder}) are satisfied. Thus, the
+information-statistical global search algorithm can be applied to the function \(\varphi(x)\) in
+order to solve the problem (\ref{eq:aux_problem}). However, \(\varphi(x)\) is defined through
+the operator \(\max\{...\}\), therefore, it is difficult to compute it directly. In
+\cite{markinStrongin1993}, a modification of the classic information-statistical
+algorithm~\cite{mixedAlg} is given, in which the values of \(\varphi(x)\) are computed
+approximately. Let us consider the modified version of this algorithm. The modification
+consists in the use of the local refinement technique described in~\cite{strOptBook}(chapter 3)
+as well as in the parallelization based on the characteristics~\cite{strOptBook}(chapter 5).
 
-Первые две итерации производятся в концевых точках \(x^0=0\) и \(x^1=1\) интервала \([0;1]\). Выбор точек \(x^{k+j}, 1\leqslant j\leqslant p\) осуществляется по правилам:
+The first two iterations are performed at the boundary points~\(x^0=0\) and \(x^1=1\) of the
+interval \([0;1]\). The choice of the points \(x^{k+j}, 1\leqslant j\leqslant p\) is performed
+according to the following rules:
 
-Step 1.
-Renumber the points in the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which includes the boundary points of the interval \([0,1]\) as well as the points of preceding trials, by the lower indices in order of increasing coordinate values  i.e.
+\textit{Step 1.} Renumber the points in the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\),
+which includes the boundary points of the interval \([0,1]\) as well as the points of preceding
+trials, by the lower indices in order of increasing coordinate values  i. e.
 \begin{displaymath}
@@
 
-Step 2. Compute the values
-\begin{equation}
-\label{step2}
-\mu_\nu=\max_{1\leqslant i\leqslant k}\dfrac{|f_\nu(x_i)-f_\nu(x_{i-1})|}{\Delta_i}, 1\leqslant \nu\leqslant m
-\end{equation}
+\textit{Step 2.} Compute the lover bound \(\mu_\nu\) for the Lipzhitz constant for each
+objective function \(f_\nu(x),1\leqslant\nu\leqslant m\):
+\begin{gather}
+\label{eq:step2_1}
+\mu_\nu=\max_{1\leqslant i\leqslant k}\dfrac{|f_\nu(x_i)-f_\nu(x_{i-1})|}{\Delta_i}, 1\leqslant
+\nu\leqslant m \\
+\label{eq:step2_2}
+\Delta_i=(x_i-x_{i-1})^\frac{1}{N}
+\end{gather}
 
-Step 3. Каждой точке \(x_i\), \(0\leqslant i\leqslant k\), сопоставить значение
+\textit{Step 3.} To each point \(x_i\), \(0\leqslant i\leqslant k\), juxtapose the value
 \begin{equation}
@@
 \end{equation}
-где
+where
 \begin{equation}
-  h(x_i,x_j)=\min\{\frac{f_\nu(x_i)-f_\nu(x_j)}{\mu_\nu}:1\leqslant \nu\leqslant m\}, 0\leqslant i,j\leqslant k
+  h(x_i,x_j)=\min\{\frac{f_\nu(x_i)-f_\nu(x_j)}{\mu_\nu}:1\leqslant \nu\leqslant m\}, 0\leqslant
+i,j\leqslant k
 \end{equation}
 
-Step 4. Для каждого интервала \((x_i,x_{i-1}),1\leqslant i\leqslant k\) вычислить величины
+\textit{Step 4.} For each interval \((x_i,x_{i-1}),1\leqslant i\leqslant k\) compute two
+characteristics:
 \begin{eqnarray}
@@
 \end{eqnarray}
-называемые характеристиками. При этом \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\), \(z^*=\min\{z_i:1\leqslant i\leqslant k\}\), а \(r>1\) и \(\alpha\in [10;30]\) --- параметры метода.
+where \(\Delta_i\) is from (\ref{eq:step2_2}), \(z^*=\min\{z_i:1\leqslant i\leqslant k\}\). \(r>1\),
+and \(\alpha\in [10;30]\) are the input parameters of the algorithm.
 
-Step 5. Если \(q\not=0\) и \(s \mod q\not=0 \), то характеристики \(R(i)\), \(1 \leqslant i \leqslant k + 1\), упорядочить в порядке убывания
+\textit{Step 5.} If \(q\not=0\) and \(s \mod q\not=0 \), then arrange the characteristics \(R(i)\),
+\(1 \leqslant i \leqslant k + 1\) in the decreasing order
 \begin{equation*}
@@
 \end{equation*}
-и выбрать \(p\) наибольших характеристик с номерами интервалов \(t_j\), \(1 \leqslant j \leqslant p\).
-Иначе то же самое сделать с характеристиками \(R^*(i),1\leqslant i\leqslant k+1\).
-Здесь \(s\) --- номер текущей итерации.
-\(q\) --- параметр метода, отвечающий за степень интенсивности локального уточнения.
-Чем меньше \(q\), тем чаще используются характеристики \(R^*\), заставляющие метод выбирать следующие точки вблизи текущего найденного минимума.
+and select \(p\) maximum characteristics with the indices of the intervals \(t_j\), \(1 \leqslant j
+\leqslant p\). Otherwise do the same with the characteristics \(R^*(i),1\leqslant i\leqslant k+1\).
+Here \(s\) is the index of current iteration and \(q\) is the parameter of the method responsible
+for the degree of intensity of the local refinement. The less \(q\), the more frequent the
+characteristics \(R^*\) are used making the method to choose the next points near the current
+minimum found.
 
-Step 6. Провести новые испытания в точках \(x^{k+j}\), \(1 \leqslant j \leqslant p\):
+\textit{Step 6.} Perform the new trials at the points \(x^{k+j}\), \(1 \leqslant j \leqslant p\):
 \begin{equation}
-  x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2} - \sign(z_{t_j} - z_{t_j-1})\frac{|z_{t_j} - z_{t_j-1}|^n}{2r}
+  x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2} - \sign(z_{t_j} - z_{t_j-1})\frac{|z_{t_j} - z_{t_j-
+1}|^n}{2r}
 \end{equation}
-Все \(p\) испытаний на этом шаге могут быть произведены параллельно на \(p\) вычислительных устройствах.
+All \(p\) trials within this step can be performed in parallel on \(p\) computer devices.
 
-The algorithm is terminated if the condition \(\Delta_{t_j}\leqslant \varepsilon\) is fulfilled at least for one of the numbers \(t_j\), \(1\leqslant j\leqslant p\); here \(\varepsilon >0\) is the predefined accuracy.
+The algorithm is terminated if the condition \(\Delta_{t_j}\leqslant \varepsilon\) is fulfilled at
+least for one of the indices \(t_j\), \(1\leqslant j\leqslant p\); here \(\varepsilon >0\) is the
+predefined accuracy.
 After the search is terminated, the set \(S(\{x^0,\dots ,x^k\})\) of all
-non-dominated points of the truncated sequence \(\{x^0,\dots ,x^k\}\) is accepted as an estimation for \(S\) from (\ref{eq:s}).
+non-dominated points of the truncated sequence \(\{x^0,\dots ,x^k\}\) is accepted as an
+estimation for \(S\) from (\ref{eq:s}).
 
-The theoretical substantiation of this method when \(p=1\) and \(q=0\) is presented in \cite{strOptBook}(chapter 3).
-Siffitient condition of convergence is: exists an iteration such that \(r\mu_\nu \geqslant 4H_\nu,1\leqslant \nu \leqslant m\).
+The theoretical substantiation of this method when \(p=1\) and \(q=0\) is presented in
+\cite{strOptBook}(chapter 3). The siffitient condition of convergence is: there exists an
+iteration such that \(r\mu_\nu \geqslant 4H_\nu, 1\leqslant \nu \leqslant m\).
 \section{Test Problems}
 \label{sec:test_problems}
-Для оценки степени ускорения сходимости модифицированного алгоритма из секции \ref{sec:algorithm} использовались следующие задачи:
+In order to evaluate the degree of speedup of the convergence of the modified algorithm from
+Section \ref{sec:algorithm}, the following problems were used:
 \begin{enumerate}
@@
     \begin{array}{l}
-      f_{1}\left(y\right) = 1 - \exp \left(-\sum_{i=1}^{n} \left(y_{i} - \frac{1}{\sqrt{n}} \right)^{2} \right) \\
-      f_{2}\left(y\right) = 1 - \exp \left(-\sum_{i=1}^{n} \left(y_{i} + \frac{1}{\sqrt{n}} \right)^{2} \right) \\
+      f_{1}\left(y\right) = 1 - \exp \left(-\sum_{i=1}^{n} \left(y_{i} - \frac{1}{\sqrt{n}}
+\right)^{2} \right) \\
+      f_{2}\left(y\right) = 1 - \exp \left(-\sum_{i=1}^{n} \left(y_{i} + \frac{1}{\sqrt{n}}
+\right)^{2} \right) \\
     \end{array}
@@
   \end{equation}
-  \item Schaffer function N. 2 \cite{Huband2006}:
+  \item
+  Viennet function
   \begin{equation}
-      \label{eq:schaffer2}
     \left \{
     \begin{array}{l}
-        f_1(y) = \begin{cases}
-                  -y,   & \text{if } y \leqslant 1 \\
-                   y-2, & \text{if } 1 < y \leqslant 3 \\
-                   4-y, & \text{if } 3 < y \leqslant 4 \\
-                   y-4, & \text{if } y > 4 \\
-                \end{cases} \\
-        f_2(y) = (y-5)^2\\
+      f_{1}\left(y\right) = 0.5(y_1^2 + y_2^2) + \sin(y_1^2 + y_2^2)\\
+      f_{2}\left(y\right) = \frac{(3y_1 - 2*y[1])}{2} + (y_1-y_2 + 1)\27 ^ 2{2} + 15\\
+      f_{3}\left(y\right) =   + \sin(y_1^2 + y_2^2)\\
     \end{array}
     \right .
-      \quad y\in [-5;10]
+    \quad y\in [-3;3]^2
   \end{equation}
+%  Schaffer function N. 2 \cite{Huband2006}:
+%    \begin{equation}
+%      \label{eq:schaffer2}
+%      \left \{
+%      \begin{array}{l}
+%        f_1(y) = \begin{cases}
+%                  -y,   & \text{if } y \leqslant 1 \\
+%                   y-2, & \text{if } 1 < y \leqslant 3 \\
+%                   4-y, & \text{if } 3 < y \leqslant 4 \\
+%                   y-4, & \text{if } y > 4 \\
+%                \end{cases} \\
+%        f_2(y) = (y-5)^2\\
+%      \end{array}
+%      \right .
+%      \quad y\in [-5;10]
+%    \end{equation}
   \item Poloni's two objective function \cite{Huband2006}:
@@
       \begin{array}{l}
-        f_{1}\left(y\right) = \left[1 + \left(A_{1} - B_{1}\left(y\right) \right)^{2} + \left(A_{2} - B_{2}\left(y\right) \right)^{2} \right] \\
+        f_{1}\left(y\right) = \left[1 + \left(A_{1} - B_{1}\left(y\right) \right)^{2} + \left(A_{2} -
+B_{2}\left(y\right) \right)^{2} \right] \\
         f_{2}\left(y\right) = \left(y_1 + 3\right)^{2} + \left(y_2 + 1 \right)^{2} \\
@@
       \begin{cases}
-        A_{1} = 0.5 \sin \left(1\right) - 2 \cos \left(1\right) + \sin \left(2\right) - 1.5 \cos \left(2\right)  \\
-        A_{2} = 1.5 \sin \left(1\right) - \cos \left(1\right) + 2 \sin \left(2\right) - 0.5 \cos \left(2\right)  \\
-        B_{1}\left(y\right) = 0.5 \sin \left(y_1\right) - 2 \cos \left(y_1\right) + \sin \left(y_2\right) - 1.5 \cos \left(y_2\right)  \\
-        B_{2}\left(y\right) = 1.5 \sin \left(y_1\right) - \cos \left(y_1\right) + 2 \sin \left(y_2\right) - 0.5 \cos \left(y_2\right)
+        A_{1} = 0.5 \sin \left(1\right) - 2 \cos \left(1\right) + \sin \left(2\right) - 1.5 \cos
+\left(2\right)  \\
+        A_{2} = 1.5 \sin \left(1\right) - \cos \left(1\right) + 2 \sin \left(2\right) - 0.5 \cos
+\left(2\right)  \\
+        B_{1}\left(y\right) = 0.5 \sin \left(y_1\right) - 2 \cos \left(y_1\right) + \sin \left(y_2\right) -
+1.5 \cos \left(y_2\right)  \\
+        B_{2}\left(y\right) = 1.5 \sin \left(y_1\right) - \cos \left(y_1\right) + 2 \sin \left(y_2\right) -
+0.5 \cos \left(y_2\right)
       \end{cases}
@@
 The computational experiments have been carried out on the Lobachevsky supercomputer at
-Lobachevsky State University of Nizhni Novgorod. A computational node included 2 Intel Sandy
+Lobachevsky State University of Nizhni Novgorod. A computational node included 2 Intel
+Sandy
 Bridge E5-2660 2.2 GHz processors, 64 Gb RAM. The CPUs had 8 cores (i. e. total 16 cores
 were available per a node).
+In the present section, we will understand the speedup of the method as the speedup in the
+number of executed iterations, not in the time of execution. If the computation of the problem
+criteria (\ref{eq:problem}) takes enough much time, the overhead costs for the execution of the
+decision rules of the optimization method are low as compared to the time of computing the
+criteria \(f_i(y), 1\leqslant i\leqslant m\).
 
-В данном разделе будем понимать под уcкорением работы метода ускорение по числу выполненных итераций, а не по времени выполнения.
-Если вычисление критериев задачи (\ref{eq:problem}) занимает достаточно много времени, то накладные затраты на реализацию решающих правил метода оптимизации невелики по сравнению с временем вычисления критериев \(f_i(y), 1\leqslant i\leqslant m\).
-
-\paragraph{Local refinement advantages.} В \cite{barkalovLebedev2016} приведены численные эксперименты, показывающие ускорение сходимости при применении техники локального уточнения в алгоритме, решающем скалярные задачи глобальной оптимизации.
-Подобных результатов стоит ожидать и от многокритериального алгоритма.
-Многокритериальный метод с локальным уточнением (MOLF) был применён к задаче Fonseca and Fleming (\ref{eq:fonseca}) при \(n=2\).
-Параметры метода были следующие: \(\varepsilon=0.01,\;r=4,\;q=4,\;\alpha=15,\;p=1\).
-До остановки метода было произведено 1176 итераций, количество найденных слабо оптимальных точек --- 90.
-При \(q=0\) (без локального уточнения) метод производит 1484 итерации, количество найденных слабо оптимальных точек --- 93.
-На рис.
-\ref{fig:fonseca_slater} представлено множество \(S\) в задаче (\ref{eq:fonseca}) и его численные оценки, полученные при \(q=0\) (fig.
-\ref{fig:fonseca_slater}.a) и \(q=4\) (fig.
-\ref{fig:fonseca_slater}.b).
+\paragraph{Local refinement advantages.} In \cite{barkalovLebedev2016} the numerical
+experiments demonstrating the speedup of convergence when using the local refinement
+technique in the algorithm solving the scalar global optimization problems are presented. One
+could expect similar results from the multi-objective algorithm as well. The multi-objective
+algorithm with local refinement (MOALR) has been applied to the  Fonseca and Fleming
+problem (\ref{eq:fonseca}) at \(n=2\). The parameters of the method were as follows:
+\(\varepsilon=0.01,\;r=4,\;q=4,\;\alpha=15,\;p=1\). Before the method stops, 1176 iterations
+have been executed, the number of found weakly optimal points was 90. At \(q=0\) (without the
+local refinement) the method had performed 1484 iterations, the number of found weakly
+optimal points was 93. In fig. \ref{fig:fonseca_slater} the set \(S\) in the problem
+(\ref{eq:fonseca}) and the numerical estimates of this one obtained at \(q=0\) (fig.
+\ref{fig:fonseca_slater}.a) and \(q=4\) (fig. \ref{fig:fonseca_slater}.b) are presented. It is
+evident also from fig. \ref{fig:fonseca_slater} that the approximate solution covers the whole
+set of the weak-optimal solutions.
 
@@
     \centering
-    \subfloat[MOLF with \(q=0\)]{{\includegraphics[width=.5\textwidth]{img/fonseca_glob.pdf} }}
-    \subfloat[MOLF with \(q=4\)]{{\includegraphics[width=.5\textwidth]{img/fonseca_loc.pdf} }}
-    \caption{Numericas estimation of \(S\) obtained by MOLF}
+    \subfloat[MOALR with
+\(q=0\)]{{\includegraphics[width=.5\textwidth]{img/fonseca_glob.pdf} }}
+    \subfloat[MOALR with
+\(q=4\)]{{\includegraphics[width=.5\textwidth]{img/fonseca_loc.pdf} }}
+    \caption{Numericas estimation of \(S\) obtained by MOALR}
     \label{fig:fonseca_slater}
@@
 
-В дальнейшем для всех экспериментов будем использовать MOLF при \(q=4\).
-\paragraph{Parallel method results.} Чтобы продемонстрировать ускорение по итерациям, которое даёт MOLF при \(p > 1\), были решены все задачи из секции \ref{sec:test_problems} при значениях \(p=1,2,4,8,16\).
-Параметры метода при этом были следующие: \(r=4.5\), \(\varepsilon=0.01\), \(\alpha=15\).
-Для задачи (\ref{eq:schaffer2}) \(\varepsilon=0.001\).
-В таблице \ref{tab:exResultsIters} приведено количество итераций (в скобках мощность оценки множества \(S\)), а в таблице \ref{tab:exResultsItersSpeedup} --- ускорение по итерациям.
+Below, MOALR with \(q=4\) was used for all experiments.
 
-Как видно из таблицы \ref{tab:exResultsIters}, мощность оценки множества \(S\) при изменении \(p\) меняется незначительно, т.
-е.
-качество полученной оценки сохраняется.
-В то же время, количество итераций соркащается пропорционально увеличению значения \(p\) (таблица \ref{tab:exResultsItersSpeedup}).
-На рис.
-\ref{fig:s_estimation_all} приведены примеры численных решений рассматриваемых задач.
+\paragraph{Parallel method results.} In order to demonstrate the speedup in iterations, which
+MOALR provides when \(p > 1\), all problems from Section \ref{sec:test_problems} have been
+solved for the values \(p=1,2,4,8,16\). The parameres of the method were the following:
+\(r=4.5\), \(\varepsilon=0.01\), \(\alpha=15\). For the problem (\ref{eq:schaffer2})
+\(\varepsilon=0.001\). The number of iterations is presented in Table \ref{tab:exResultsIters}
+(in the braces, the cardinalityof the set \(S\) is given) and the speedup in iterations is presented
+in in Table \ref{tab:exResultsItersSpeedup}. As it is evident form Table
+\ref{tab:exResultsIters}, the cardinality of the set \(S\) changed insuffiliently when varying \(p\)
+i. e. the quality of the obtained estimates remained the same. At the same time, the number of
+iterations decreased linearly with increasing value of \(p\) (Table
+\ref{tab:exResultsItersSpeedup}). In fig. \ref{fig:s_estimation_all}, the examples of the
+numerical solutions of the considered problems are given.
+
 \begin{table}
   \centering
-  \caption{Results of numerical experiments: number of iterations iterations}
+  \caption{Results of numerical experiments: number of iterations}
   \label{tab:exResultsIters}
@@
 \hline
-Schaffer function N. 2 & 271(158) & 136(163) & 64(146) & 32(142) & 16(141)\\
+Viennet problem & 4896(276) & 2156(273) & 1226(270) & 631(287) & 286(274)\\
 \hline
@@
 \hline
-Schaffer problem N. 2 & 1.99 & 4.23 & 8.47 & 16.94\\
+Viennet problem & 2.27 & 3.99 & 7.76 & 17.12\\
 \hline
@@
     \centering
-    \subfloat[Markin-Strongin problem]{{\includegraphics[width=.5\textwidth]{img/strongin.pdf} }}
-    \subfloat[Fonseca and Fleming 3d problem]{{\includegraphics[width=.5\textwidth]{img/fonseca.pdf} }}
+    \subfloat[Markin-Strongin
+problem]{{\includegraphics[width=.5\textwidth]{img/strongin.pdf} }}
+    \subfloat[Fonseca and Fleming 3d
+problem]{{\includegraphics[width=.5\textwidth]{img/fonseca.pdf} }}
 
-    \subfloat[Schaffer problem N. 2]{{\includegraphics[width=.5\textwidth]{img/schaffer2.pdf} }}
+    \subfloat[Viennet problem]{{\includegraphics[width=.5\textwidth]{img/viennet.png}
+}}
     \subfloat[Poloni's problem]{{\includegraphics[width=.5\textwidth]{img/poloni.pdf} }}
-    \caption{Numericas estimation of \(S\) obtained by MOLF}
+    \caption{Numericas estimations of \(S\) obtained by MOALR}
     \label{fig:s_estimation_all}
@@
 
-\paragraph{Parallel method time speedup.} Целевые функции всех перечисленных в разделе задачи обладают невысокой вычислительной сложностью.
-Чтобы продемонстрировать возможность получения ускорения по времени для задач с трудновычислимыми критериями, в критерии были внесены дополнительные интенсивные вычисления с плавающей точкой, не влияющие на результирующие значения.
-Полученные при этом ускорения приведены в таблице \ref{tab:exResultsTimeSpeedup}.
-Ускорение по времени во всех случаях меньше, чем ускорение по итерациям, т.к.
-часть времени выполнения занимает работа метоад оптимизации, однако распареллеливание в большинстве случаев довольно эффективно и в случае 16 рабочих потоков.
-При ещё большей вычислительной сложности критериев стоит ожидать увеличения ускорения по времени для большого числа потоков.
+\paragraph{Parallel method time speedup.} The objective functions of all problems listed in this
+section are featured by low computational complexity. In order to demonstrate the possibility to
+obtain a speedup in time for the problems with the hard-to-compute criteria, additinal intensive
+froating point computations not affecting the resulting values were introduced into the criteria.
+The speedups obtained are given in Table \ref{tab:exResultsTimeSpeedup}. In all cases, the
+speedup in time was less than the speedip in iterations because the operation of the optimizaion
+method itself takes a part of the execution time. However, in most cases the parallelizaion was
+efficient enough even in the case of 16 CPU threads. For more computation costly criteria, one can
+expect the increasing of the speedup in time for a large number of streams.
 
@@
 \hline
-Schaffer problem N. 2 & 27.21 & 1.91 & 3.79 & 6.39 & 6.37\\
+Viennet problem & - & 1.91 & 3.79 & 6.39 & 6.37\\
 \hline
@@
 \section{Conclusion}
-В работе был рассмотрен метод решения многокритериальных задач оптимизации, позволяющий находить равномерную оценку множества слабо-оптимальных точек.
-К нему применены техники распараллеливания и ускорения сходимости, общие для всех алгоритмов подобного структуры \cite{strOptBook}, chapter 3.
-Проведённые численные эксперименты показали наличие ускорения сходимости при использовании local refinement.
-Также для данного метода оказалось эффективным распареллеливание по характеристикам.
-Как и для информационных методов скалярной оптимизации, ускорение по итерациям при распареллеливании по характеристикам оказывается линейным в большинстве случаев \cite{barkalovLebedef2016}.
-Исходя из результатов эскпериментов, параллельный MOLF применим в задачах низкой размерности с трудновычислимыми критериями.
+In the present work, a method for solving the multi-objective optimizaion problems allowing
+finding a uniform estimate of the set of the weakly-optimal points has been considered. The
+techniques of parallelization and of the acceleration of convergence common for all algorithms
+of similar structure .have been applied. The numerical experiments performed have
+demonstrated a speedup in convergence when using the local refinement. Also, the
+parallelization based on the characteristics was found to be efficient for this method. The
+speedup in iterations at the parallelization based on the characteristics appeared to be linear in
+most cases as for the information methods of scalar optimization (see the results of the
+parlallelization of the scalar optimization method, for example, in
+\cite{barkalovLebedef2016}). From the results of the experiments, one can conclude that it is
+expedient to apply sthe parallel MOALR in the problems of low dimensionality (1 to 3) with
+the computation-costly criteria. Besides, the property of uniform convergence of the considered
+algorithm to the whole set of the weakly-optimal solutions has been demonstrated on the
+numerical  examples.
 %

### 2017-07-10 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: tiny fixes

@@
 considered approach is based on the application of the information-statistical algorithm to some
-reduced unicriterial problem, the set of the global optima in which coincides to the set of the
+reduced scalar problem, the set of the global optima in which coincides to the set of the
 weakly efficient solutions of the initial multi-objective problem. The sequential version of this
 method has been considered earlier. In the present work, the parallelization scheme based on
-the characteristics, which is common for all information-statistical global search algorithms,
+the characteristics, which is common for all information global search algorithms,
 was applied to the sequential algorithm of the multi-objective optimization. Also, in the present
@@
 problems~\cite{GergelKozinov2017}, various approaches to the solving the problems with the
-convex~ \cite{ZhangZuo2013} or linear~\cite{Benson1998} criteria. The main problem in the
+convex~\cite{ZhangZuo2013} or linear~\cite{Benson1998} objectives. The main problem in the
 using of majority of these methods is the absence of the theoretical guarantee of the uniform
@@
 As it has been shown in~\cite{markinStrongin1993}, the set of weakly efficient solutions of the
-reduced problem (\(\ref{eq:problem}\)) coincides to the set of the globally optimal solutions of
+reduced problem (\ref{eq:problem}) coincides to the set of the globally optimal solutions of
 the problem (\ref{eq:aux_problem}) i. e.
@@
 the operator \(\max\{...\}\), therefore, it is difficult to compute it directly. In
-\cite{markinStrongin1993}, a modification of the classic information-statistical
-algorithm~\cite{mixedAlg} is given, in which the values of \(\varphi(x)\) are computed
+\cite{markinStrongin1993}, a modification of the classic information
+global search algorithm~\cite{mixedAlg} is given, in which the values of \(\varphi(x)\) are computed
 approximately. Let us consider the modified version of this algorithm. The modification
@@
 \end{equation}
-All \(p\) trials within this step can be performed in parallel on \(p\) computer devices.
+All \(p\) trials within this step can be performed in parallel on \(p\) computing devices.
 
@@
       f_{1}\left(y\right) = 0.5(y_1^2 + y_2^2) + \sin(y_1^2 + y_2^2)\\
-      f_{2}\left(y\right) = \frac{(3y_1 - 2*y[1])}{2} + (y_1-y_2 + 1)\27 ^ 2{2} + 15\\
-      f_{3}\left(y\right) =   + \sin(y_1^2 + y_2^2)\\
+      f_{2}\left(y\right) = \frac{(3y_1 - 2y_2 + 4)^2}{8} + \frac{(y_1-y_2 + 1)^2}{27} + 15\\
+      f_{3}\left(y\right) = \frac{1}{y_1^2+y_2^2 + 1} -1.1 \exp\{-(y_1^2+y_2^2)\}\\
     \end{array}
@@
 Lobachevsky State University of Nizhni Novgorod. A computational node included 2 Intel
-Sandy
-Bridge E5-2660 2.2 GHz processors, 64 Gb RAM. The CPUs had 8 cores (i. e. total 16 cores
+Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM. The CPUs had 8 cores (i. e. total 16 cores
 were available per a node).
@@
 solved for the values \(p=1,2,4,8,16\). The parameres of the method were the following:
-\(r=4.5\), \(\varepsilon=0.01\), \(\alpha=15\). For the problem (\ref{eq:schaffer2})
-\(\varepsilon=0.001\). The number of iterations is presented in Table \ref{tab:exResultsIters}
+\(r=4.5\), \(\varepsilon=0.01\), \(\alpha=15\). The number of iterations is presented in Table \ref{tab:exResultsIters}
 (in the braces, the cardinalityof the set \(S\) is given) and the speedup in iterations is presented
@@
 efficient enough even in the case of 16 CPU threads. For more computation costly criteria, one can
-expect the increasing of the speedup in time for a large number of streams.
+expect the increasing of the speedup in time for a large number of threads.
 
@@
 \hline
-Viennet problem & - & 1.91 & 3.79 & 6.39 & 6.37\\
+Viennet problem & TBD & TBD & TBD & TBD & TBD\\
 \hline
@@
 techniques of parallelization and of the acceleration of convergence common for all algorithms
-of similar structure .have been applied. The numerical experiments performed have
+of similar structure have been applied. The numerical experiments performed have
 demonstrated a speedup in convergence when using the local refinement. Also, the

### 2017-07-11 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: update time speedup table
u_pdc: fix various typos

@@
 \end{equation}
-
 which is usually referred as the set of semi-efficient (or weakly efficient) solutions. The
@@
 \end{eqnarray}
-where \(\Delta_i\) is from (\ref{eq:step2_2}), \(z^*=\min\{z_i:1\leqslant i\leqslant k\}\). \(r>1\),
+where \(\Delta_i\) is from (\ref{eq:step2_2}), \(z^*=\min\{z_i:1\leqslant i\leqslant k\}\), \(r>1\),
 and \(\alpha\in [10;30]\) are the input parameters of the algorithm.
@@
   \item
-  Viennet function
+  Viennet function~\cite{Huband2006}:
   \begin{equation}
@@
 The computational experiments have been carried out on the Lobachevsky supercomputer at
-Lobachevsky State University of Nizhni Novgorod. A computational node included 2 Intel
+State University of Nizhny Novgorod. A computational node included 2 Intel
 Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM. The CPUs had 8 cores (i. e. total 16 cores
@@
 local refinement) the method had performed 1484 iterations, the number of found weakly
-optimal points was 93. In fig. \ref{fig:fonseca_slater} the set \(S\) in the problem
+optimal points was 93. In Fig. \ref{fig:fonseca_slater} the set \(S\) in the problem
 (\ref{eq:fonseca}) and the numerical estimates of this one obtained at \(q=0\) (fig.
@@
 \(r=4.5\), \(\varepsilon=0.01\), \(\alpha=15\). The number of iterations is presented in Table \ref{tab:exResultsIters}
-(in the braces, the cardinalityof the set \(S\) is given) and the speedup in iterations is presented
+(in the braces, the cardinality of the set \(S\) is given) and the speedup in iterations is presented
 in in Table \ref{tab:exResultsItersSpeedup}. As it is evident form Table
-\ref{tab:exResultsIters}, the cardinality of the set \(S\) changed insuffiliently when varying \(p\)
+\ref{tab:exResultsIters}, the cardinality of the weak-optimal solutions set changed insufficiently when varying \(p\)
 i. e. the quality of the obtained estimates remained the same. At the same time, the number of
 iterations decreased linearly with increasing value of \(p\) (Table
-\ref{tab:exResultsItersSpeedup}). In fig. \ref{fig:s_estimation_all}, the examples of the
+\ref{tab:exResultsItersSpeedup}). In Fig. \ref{fig:s_estimation_all}, the examples of the
 numerical solutions of the considered problems are given.
@@
 The speedups obtained are given in Table \ref{tab:exResultsTimeSpeedup}. In all cases, the
-speedup in time was less than the speedip in iterations because the operation of the optimizaion
+speedup in time was less than the speedup in iterations because the operation of the optimizaion
 method itself takes a part of the execution time. However, in most cases the parallelizaion was
@@
 \hline
-Viennet problem & TBD & TBD & TBD & TBD & TBD\\
+Viennet problem & 1488.6 & 2.22 & 3.64 & 6.98 & 13.49\\
 \hline
@@
 \cite{barkalovLebedef2016}). From the results of the experiments, one can conclude that it is
-expedient to apply sthe parallel MOALR in the problems of low dimensionality (1 to 3) with
+expedient to apply the parallel MOALR in the problems of low dimensionality (1 to 3) with
 the computation-costly criteria. Besides, the property of uniform convergence of the considered

### 2017-07-12 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: small fix in intro section

@@
 convex~\cite{ZhangZuo2013} or linear~\cite{Benson1998} objectives. The main problem in the
-using of majority of these methods is the absence of the theoretical guarantee of the uniform
+using of the most of these methods is the absence of the theoretical guarantee of the uniform
 convergence to the whole sought set of the Pareto-optimal solutions on a wide class of

### 2017-07-13 2017/URAL-PDC/u_pdc_paper.tex

@@
 \begin{abstract}
-In the present work, a parallel algorithm for the multi-objective optimization is considered. The
-considered approach is based on the application of the information-statistical algorithm to some
+This work considers a parallel algorithm for the multi-objective optimization. The
+algorithm is based on the application of the information-statistical approach for solving
 reduced scalar problem, the set of the global optima in which coincides to the set of the
-weakly efficient solutions of the initial multi-objective problem. The sequential version of this
-method has been considered earlier. In the present work, the parallelization scheme based on
-the characteristics, which is common for all information global search algorithms,
-was applied to the sequential algorithm of the multi-objective optimization. Also, in the present
-work one of the techniques of accounting for the local properties of the optimized function
-allowing accelerating the convergence essentially has been considered for the multi-objective
-method for the first time.
+weakly efficient solutions of the initial multi-objective problem. In the present work the parallelization scheme, which is common for all information-statistical global search algorithms,
+was applied to the sequential algorithm of the multi-objective optimization. Also, in this
+work one of the techniques of accounting for the local properties of the objective function
+allowing accelerating the convergence essentially has been applied to the multi-objective
+problems.
 
@@
   \min\{f(y): y\in D\}, D=\{y\in \mathbb{R}^n: a_i \leqslant y_i \leqslant b_i, 1\leqslant i
-\leqslant n \}
+\leqslant n \}.
 \end{equation}
 Let us assume the components of the vector function (the partial criteria) \(f_i(y), 1\leqslant
-i\leqslant m\) to satisfy the Lipschitz condition with the Lipschitz constant \(L_i\) in \(D\):
+i\leqslant m\), to satisfy the Lipschitz condition with the Lipschitz constant \(L_i\) in \(D\):
 \begin{displaymath}
@@
 
-The set \(S(D)\in D\) of strictly non-dominated points from the search domain is accepted as the
-solution of the problem (\ref{eq:problem}) usually i. e.
+The set \(S(D)\in D\) of strictly non-dominated points from the search domain is usually accepted as the
+solution of the problem (\ref{eq:problem}), i.e.
 
@@
   \label{eq:slater}
-  S(D) = \{y\in D: \nexists z\in D, f_i(z)<f_i(y),1\leqslant i \leqslant m\}
+  S(D) = \{y\in D: \nexists z\in D, f_i(z)<f_i(y),1\leqslant i \leqslant m\},
 \end{equation}
@@
 
-The use of the evolvents \(y(x)\) i. e. the space-filling curves is a classic dimensionality
-reduction scheme for global optimization algorithms~\cite{evolvents2013}.
+The use of the space-filling curve (\textit{evolvent}) \(y(x)\)
 \begin{displaymath}
@@
 \end{displaymath}
-\par
+is a classic dimensionality reduction scheme for global optimization algorithms~\cite{evolvents2013}.
 Such a mapping allows the reduction of the problem (\(\ref{eq:problem}\)) stated in a
@@
 \label{eq:holder}
-|f_i(y(x_1))-f_i(y(x_2))|\leqslant H_i{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
+|f_i(y(x_1))-f_i(y(x_2))|\leqslant H_i{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1],
 \end{equation}
@@
 \begin{displaymath}
-H_i=4L_id\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant n\}
+H_i=4L_id\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant n\}.
 \end{displaymath}
@@
 Therefore, not limiting the generality, one can consider the solving of the
-one-dimensional problem \(\min\{f(y(x)): x\in [0;1]\}\), satisfying the H\"older condition. The
+one-dimensional problem \(\min\{f(y(x)): x\in [0,1]\}\), satisfying the H\"older condition. The
 issues of the numerical building the mapping like a Peano curve and the corresponding theory
@@
 built numerically is an approximation of the theoretical Peano curve with a precision of the
-order \(2^{-m}\) where \(m\) is the building parameter of the evolvent.
+order \(2^{-m}\), where \(m\) is the building parameter of the evolvent.
 
@@
 \begin{equation}
-  \varphi(x)=\max\{h(x,y):y\in [0;1]\},x\in [0;1].
+  \varphi(x)=\max\{h(x,y):y\in [0,1]\},x\in [0,1].
 \end{equation}
@@
   \label{eq:aux_problem}
-  \varphi^*=\min\{\varphi(x):x\in [0;1]\}.
+  \varphi^*=\min\{\varphi(x):x\in [0,1]\}.
 \end{equation}
@@
 reduced problem (\ref{eq:problem}) coincides to the set of the globally optimal solutions of
-the problem (\ref{eq:aux_problem}) i. e.
+the problem (\ref{eq:aux_problem}), i.e.
 \begin{equation}
   \label{eq:s}
-  S([0;1])=\{x\in [0;1]:\varphi(x)=\varphi^*\}
+  S([0,1])=\{x\in [0,1]:\varphi(x)=\varphi^*\}.
 \end{equation}
@@
 the operator \(\max\{...\}\), therefore, it is difficult to compute it directly. In
-\cite{markinStrongin1993}, a modification of the classic information
+\cite{markinStrongin1993}, a modification of the original information
 global search algorithm~\cite{mixedAlg} is given, in which the values of \(\varphi(x)\) are computed
-approximately. Let us consider the modified version of this algorithm. The modification
+approximately. Let us consider a new version of this algorithm. The modification
 consists in the use of the local refinement technique described in~\cite{strOptBook}(chapter 3)
@@
 The first two iterations are performed at the boundary points~\(x^0=0\) and \(x^1=1\) of the
-interval \([0;1]\). The choice of the points \(x^{k+j}, 1\leqslant j\leqslant p\) is performed
+interval \([0,1]\). The choice of the points \(x^{k+j}, 1\leqslant j\leqslant p\), is performed
 according to the following rules:
@@
 which includes the boundary points of the interval \([0,1]\) as well as the points of preceding
-trials, by the lower indices in order of increasing coordinate values  i. e.
+trials, by the lower indices in order of increasing coordinate values, i.e.
 \begin{displaymath}
-  0=x_0<x_1<\dotsc<x_{k+1}=1
+  0=x_0<x_1<\dotsc<x_{k+1}=1.
 \end{displaymath}
 
-\textit{Step 2.} Compute the lover bound \(\mu_\nu\) for the Lipzhitz constant for each
+\textit{Step 2.} Compute the lower bound \(\mu_\nu\) for the Lipschitz constant for each
 objective function \(f_\nu(x),1\leqslant\nu\leqslant m\):
@@
 \mu_\nu=\max_{1\leqslant i\leqslant k}\dfrac{|f_\nu(x_i)-f_\nu(x_{i-1})|}{\Delta_i}, 1\leqslant
-\nu\leqslant m \\
+\nu\leqslant m, \\
 \label{eq:step2_2}
-\Delta_i=(x_i-x_{i-1})^\frac{1}{N}
+\Delta_i=(x_i-x_{i-1})^\frac{1}{N}.
 \end{gather}
@@
 \begin{equation}
-  h(x_i,x_j)=\min\{\frac{f_\nu(x_i)-f_\nu(x_j)}{\mu_\nu}:1\leqslant \nu\leqslant m\}, 0\leqslant
+  h(x_i,x_j)=\min \left\{ \frac{f_\nu(x_i)-f_\nu(x_j)}{\mu_\nu}:1\leqslant \nu\leqslant m \right\}, 0\leqslant
 i,j\leqslant k
@@
 
-\textit{Step 4.} For each interval \((x_i,x_{i-1}),1\leqslant i\leqslant k\) compute two
+\textit{Step 4.} For each interval \((x_i,x_{i-1}),1\leqslant i\leqslant k\), compute two
 characteristics:
 \begin{eqnarray}
-  R(i) = \Delta_i + \frac{(z_i-z_{i-1})^2}{r^2\Delta_i}-\frac{z_i+z_{i-1}}{2r} \\
+  R(i) = \Delta_i + \frac{(z_i-z_{i-1})^2}{r^2\Delta_i}-\frac{z_i+z_{i-1}}{2r}, \\
   R^*(i)=\frac{R(i)}{\sqrt{(z_i-z^*)(z_{i-1}-z^*)} + 1.5^{-\alpha}},
 \end{eqnarray}
-where \(\Delta_i\) is from (\ref{eq:step2_2}), \(z^*=\min\{z_i:1\leqslant i\leqslant k\}\), \(r>1\),
-and \(\alpha\in [10;30]\) are the input parameters of the algorithm.
+where \(\Delta_i\) is from (\ref{eq:step2_2}), \(z^*=\min\{z_i:1\leqslant i\leqslant k\}\); \(r>1\) and
+\(\alpha\in [10,30]\) are the input parameters of the algorithm.
 
@@
 \leqslant p\). Otherwise do the same with the characteristics \(R^*(i),1\leqslant i\leqslant k+1\).
-Here \(s\) is the index of current iteration and \(q\) is the parameter of the method responsible
+Here \(s\) is the index of current iteration and integer \(q\) is the parameter of the method responsible
 for the degree of intensity of the local refinement. The less \(q\), the more frequent the
@@
 The theoretical substantiation of this method when \(p=1\) and \(q=0\) is presented in
-\cite{strOptBook}(chapter 3). The siffitient condition of convergence is: there exists an
-iteration such that \(r\mu_\nu \geqslant 4H_\nu, 1\leqslant \nu \leqslant m\).
+\cite{strOptBook}(chapter 3). 
+%The siffitient condition of convergence is: there exists an iteration such that \(r\mu_\nu \geqslant 4H_\nu, 1\leqslant \nu \leqslant m\).
 \section{Test Problems}
@@
 \begin{enumerate}
-  \item Markin-Strongin problem from \cite{markinStrongin1993}:
+  \item Markin-Strongin problem \cite{markinStrongin1993}:
     \begin{equation}
@@
 %      \right .
-%      \quad y\in [0;1]^2
+%      \quad y\in [0,1]^2
 %    \end{equation}
@@
 In the present section, we will understand the speedup of the method as the speedup in the
-number of executed iterations, not in the time of execution. If the computation of the problem
-criteria (\ref{eq:problem}) takes enough much time, the overhead costs for the execution of the
+number of executed iterations, not in the time of execution. If the problem
+criteria (\ref{eq:problem}) is time-consuming, the overhead costs for the execution of the
 decision rules of the optimization method are low as compared to the time of computing the
@@
 MOALR provides when \(p > 1\), all problems from Section \ref{sec:test_problems} have been
-solved for the values \(p=1,2,4,8,16\). The parameres of the method were the following:
+solved for the values \(p=1,2,4,8,16\). The parameters of the method were the following:
 \(r=4.5\), \(\varepsilon=0.01\), \(\alpha=15\). The number of iterations is presented in Table \ref{tab:exResultsIters}
@@
 section are featured by low computational complexity. In order to demonstrate the possibility to
-obtain a speedup in time for the problems with the hard-to-compute criteria, additinal intensive
-froating point computations not affecting the resulting values were introduced into the criteria.
+obtain a speedup in time for the problems with the time-consuming criteria, additional intensive
+floating point computations not affecting the resulting values were introduced into the criteria.
 The speedups obtained are given in Table \ref{tab:exResultsTimeSpeedup}. In all cases, the
-speedup in time was less than the speedup in iterations because the operation of the optimizaion
-method itself takes a part of the execution time. However, in most cases the parallelizaion was
+speedup in time was less than the speedup in iterations because the operation of the optimization
+method itself takes a part of the execution time. However, in most cases the parallelization was
 efficient enough even in the case of 16 CPU threads. For more computation costly criteria, one can
@@
 \section{Conclusion}
-In the present work, a method for solving the multi-objective optimizaion problems allowing
+In the present work, a method for solving the multi-objective optimization problems allowing
 finding a uniform estimate of the set of the weakly-optimal points has been considered. The
@@
 most cases as for the information methods of scalar optimization (see the results of the
-parlallelization of the scalar optimization method, for example, in
+parallelization of the scalar optimization method, for example, in
 \cite{barkalovLebedef2016}). From the results of the experiments, one can conclude that it is
 expedient to apply the parallel MOALR in the problems of low dimensionality (1 to 3) with
-the computation-costly criteria. Besides, the property of uniform convergence of the considered
+the computation-costly criteria. Moreover, the property of uniform convergence of the considered
 algorithm to the whole set of the weakly-optimal solutions has been demonstrated on the

### 2017-07-13 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: change title, add acknowledgments section

@@
 %
-\title{Parallel Multi-objective Optimization on CPU Using Information Framework for
+\title{Parallel Multi-objective Optimization Using Information Framework for
 Constructing Global Optimization Algorithms}
@@
 numerical examples.
+
+\section*{Acknowledgments}
+The study was supported by the Russian Science Foundation, project No 16-11-10150.
 %
@@
 %
-%\bibliographystyle{plain}
-%\bibliographystyle{alpha}
-\bibliographystyle{unsrt}
-%\bibliographystyle{splncs03}
+\bibliographystyle{splncs03}
 \bibliography{u_pdc_refs}

### 2017-07-13 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: MOALR->MOAR

@@
 could expect similar results from the multi-objective algorithm as well. The multi-objective
-algorithm with local refinement (MOALR) has been applied to the  Fonseca and Fleming
+algorithm with local refinement (MOAR) has been applied to the  Fonseca and Fleming
 problem (\ref{eq:fonseca}) at \(n=2\). The parameters of the method were as follows:
@@
     \centering
-    \subfloat[MOALR with
+    \subfloat[MOAR with
 \(q=0\)]{{\includegraphics[width=.5\textwidth]{img/fonseca_glob.pdf} }}
-    \subfloat[MOALR with
+    \subfloat[MOAR with
 \(q=4\)]{{\includegraphics[width=.5\textwidth]{img/fonseca_loc.pdf} }}
-    \caption{Numericas estimation of \(S\) obtained by MOALR}
+    \caption{Numericas estimation of \(S\) obtained by MOAR}
     \label{fig:fonseca_slater}
@@
 
-Below, MOALR with \(q=4\) was used for all experiments.
+Below, MOAR with \(q=4\) was used for all experiments.
 
 \paragraph{Parallel method results.} In order to demonstrate the speedup in iterations, which
-MOALR provides when \(p > 1\), all problems from Section \ref{sec:test_problems} have been
+MOAR provides when \(p > 1\), all problems from Section \ref{sec:test_problems} have been
 solved for the values \(p=1,2,4,8,16\). The parameters of the method were the following:
@@
     \subfloat[Poloni's problem]{{\includegraphics[width=.5\textwidth]{img/poloni.pdf} }}
-    \caption{Numericas estimations of \(S\) obtained by MOALR}
+    \caption{Numericas estimations of \(S\) obtained by MOAR}
     \label{fig:s_estimation_all}
@@
 \cite{barkalovLebedef2016}). From the results of the experiments, one can conclude that it is
-expedient to apply the parallel MOALR in the problems of low dimensionality (1 to 3) with
+expedient to apply the parallel MOAR in the problems of low dimensionality (1 to 3) with
 the computation-costly criteria. Moreover, the property of uniform convergence of the considered

### 2017-07-14 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: fix missing h(x,y) definition

@@
 presented in~\cite{markinStrongin1993}. Let
-\begin{equation}
-  \varphi(x)=\max\{h(x,y):y\in [0,1]\},x\in [0,1].
-\end{equation}
+\begin{eqnarray*}
+  \varphi(x)=\max\{h(x,y):y\in [0,1]\},x\in [0,1],\\
+  h(x,y)=\max\{f_i(x)-f_i(y):1\leqslant i\leqslant\nu\}.
+\end{eqnarray*}
 Let us consider a scalar problem
@@
 optimal points was 93. In Fig. \ref{fig:fonseca_slater} the set \(S\) in the problem
-(\ref{eq:fonseca}) and the numerical estimates of this one obtained at \(q=0\) (fig.
-\ref{fig:fonseca_slater}.a) and \(q=4\) (fig. \ref{fig:fonseca_slater}.b) are presented. It is
-evident also from fig. \ref{fig:fonseca_slater} that the approximate solution covers the whole
+(\ref{eq:fonseca}) and the numerical estimates of this one obtained at \(q=0\) (Fig.
+\ref{fig:fonseca_slater}.a) and \(q=4\) (Fig. \ref{fig:fonseca_slater}.b) are presented. It is
+evident also from Fig. \ref{fig:fonseca_slater} that the approximate solution covers the whole
 set of the weak-optimal solutions.

### 2017-07-27 2017/URAL-PDC/u_pdc_paper.tex

@@
 %
-\title{Parallel Multi-objective Optimization Using Information Framework for
-Constructing Global Optimization Algorithms}
+\title{Parallel Multi-objective Optimization Method for Finding Complete Set of Weakly Efficient Solutions
+%Using Information Framework for Constructing Global Optimization Algorithms
+}
 %
@@
 %
-\institute{State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
+\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia\\
 \email{sovrasov.vlad@gmail.com}}
@@
   \varphi(x)=\max\{h(x,y):y\in [0,1]\},x\in [0,1],\\
-  h(x,y)=\max\{f_i(x)-f_i(y):1\leqslant i\leqslant\nu\}.
+  h(x,y)=\max\{f_i(x)-f_i(y):1\leqslant i\leqslant m\}.
 \end{eqnarray*}
@@
 as well as in the parallelization based on the characteristics~\cite{strOptBook}(chapter 5).
+
+The algorithm for solving problem (\ref{eq:problem}) involves constructing a sequence of points $x^i$, where the values of the objective functions are calculated.
+Let us call the functions values calculation process (including construction of image $y^i=y(x^i)$) the trial, and the tuple $\{x^i, y^i , f_1(y^i), ...
+f_m(y^i)\}$, the trial result.
+At each iteration $p$ trials is carried out in parallel, and the set of tuples
+\[ 
+\{x^i, y^i , f_1(y^i), ... f_m(y^i)\}, 1 \leq i \leq k=np,
+\] 
+makes up the search data collected using the method after carrying out $n$ steps. The rules that define the work of a parallel multi-objective algorithm are as follows.
 

### 2017-08-22 2017/JOGO/Barkalov_Strongin.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 Стронгин (JOGO)\Review\JOGO

\RequirePackage{fix-cm}
%

%\documentclass{svjour3}                     % onecolumn (standard format)
\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)

%надо закомментировать следующую строку!
%\usepackage[top=3cm, bottom=3cm, right=5cm]{geometry}

%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Journal of Global Optimization}
%
\begin{document}

%надо закомментировать следующую строку!
%\fontsize{12}{13pt}\selectfont

\title{Solving a set of global optimization problems by the parallel technique with uniform convergence
%On the uniform convergance of Global Search Algorithm when solving in parallel a series of optimization problems
\thanks{This study was supported by the Russian Science Foundation, project No 16-11-10150.}
}

\titlerunning{Solving a set of global optimization problems}        % if too long for running head

\author{Konstantin Barkalov         \and
        Roman Strongin 
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{K. Barkalov  \at
               Institute of Information Technology, Mathematics and Mechanics\\ Lobachevsky State University of Nizhni Novgorod\\ Nizhni Novgorod, Russia \\
              \email{konstantin.barkalov@itmm.unn.ru} 
           \and
           R. Strongin \at
					Institute of Information Technology, Mathematics and Mechanics\\ Lobachevsky State University of Nizhni Novgorod\\ Nizhni Novgorod, Russia \\
							\email{strongin@unn.ru} 
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}

In this paper, we consider solving a set of global optimization problems in parallel. The proposed novel algorithm provides uniform convergence to the set of solutions for all problems treated simultaneously. The current accuracy for each particular solution is estimated by the difference in each coordinate from the point of global decision. The main statement is given in the corresponding theorem. For the sake of illustration some computational results with hundreds of multidimensional global problems are provided.

%This paper considers solving a series of Lipschitz optimization problems in parallel. A novel algorithm has been proposed, which provides the convergence to the solutions to all problems in the series simultaneously, regardless of the particular form of the objective function in various problems. A uniform decrease of the distance between the optimizer and current estimate of this one in all problems is the main advantage of the proposed algorithm. A theorem on the uniform convergence of the algorithm has been proven. The results of the computational experiments carried out on several hundred global optimization problems of various dimensions are presented.

\keywords{ global optimization \and Peano-type space filling curves \and uniform convergence \and parallel computing}

\end{abstract}

\section{Introduction} \label{intro}

This paper considers global optimization problems of the following form:
\begin{eqnarray}\label{main_problem}
& \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
& D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
\end{eqnarray}
The objective function is assumed to satisfy the Lipschitz condition 
\[
\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in D,\; 0<L<\infty,
\]
with the constant $L$ unknown a priori. The analytical form of the function may be unknown, i.e. the objective function may be defined by an algorithm for computing the function values within the search domain (a ``black-box'' function). At the same time, it is assumed that even a single computation of the objective function value may be a time-consuming operation since it involves numerical modeling when solving applied problems.

The assumption regarding the fulfillment of the Lipschitz condition for the objective function is typical of many approaches (see, for example \cite{Strongin2000}--\cite{Paulavicius2014}). In recent years, the theory and the methods for parallel Lipschitz global optimizations have been extensively developed. Parallel implementations have been proposed for many well-known algorithms \cite{Gergel2005}--\cite{Paulavicius2011}.

Usually, the methods of global optimization (both sequential and parallel) are intended to solve a single optimization problem. If one needs to solve a series of $q$ problems, the problems in the series are solved sequentially, one after another. Therefore, the optimum estimation in the $i$-th problem in the series remains undefined until all preceding problems of the series (with the indices $1, 2, ..., i-1$) have been completely solved. In the case of limited computational resources, the optima estimates in the problems $i+1,..., q$ will not be obtained if the computation resources are exhausted while solving the $i$-th problem.

Situations when a series of $q$ problems must be solved are not an extraordinary. For example, a series of scalar problems arises when one needs to find a Pareto set in solving multi-objective optimization problems. In this case, the solution of a single scalar problem corresponds to one of the Pareto optimal points of a multi-objective problem (for reference, a review of the scalarization techniques is presented in \cite{Ehrgott2005}). A series of optimization problems also arises when using dimension reduction methods to solve multidimensional optimization problems. It is known that a multidimensional problem can be reduced to a set of problems of lower dimension. A detailed description of various dimension reduction schemes is presented in \cite{Strongin2000,Sergeyev2013}. 
%A series of test problems can be obtained using some generator (see, for example, \cite{Gaviano2003}).

When solving a set of problems, it is useful to get the estimates of the optima for all the problems at once as early as possible. This may be necessary to evaluate the expedience of continuing the search process at any point in time. It is also desirable to have the estimates of optima for all problems with the same precision. Running several independent processes, each solving one of the problems in the series on a parallel computer system, has a number of disadvantages. First, a workload imbalance between the processors will occur. If solving the $i$-th problem requires considerably fewer iterations of the method than solving the $j$-th problem, the processor tasked with handling the $i$-th problem would remain idle after completing the task. Second, the estimates of the optima will be obtained with different precision in different problems. Simpler problems will be solved with higher precision, whereas precision will be lower for more complex problems.

The goal of this study was to develop a method for solving a series of global optimization problems, which would be free of the drawbacks listed above. Namely, the method should ensure: 
\begin{itemize}
	\item a uniform load for all processors/cores employed;
	\item a uniform convergence to the solutions of all problems in the series.
\end{itemize} 
%new text begins
Here and below uniform convergence is taken to mean proportional decreasing of the distance between the best iteration point and the global optimizer in all the problems in the series.
%new text ends

The proposed method is a further development of the Global Search Algorithm (GSA) and of its parallel generalization described in detail in \cite{Sergeyev1994}--\cite{Barkalov2016} as well as in monographs \cite{Strongin2000,Sergeyev2013}.
%new text begins
GSA was developed in the framework of a stochastic model as a one-step optimal Bayesian method that distinguishes it from the methods optimal in the worst case (see, for example, \cite{Zilinskas2013}). The idea of average optimality of global optimization algorithms rests on the assumption (which is typical of applied problems) that the worst case objective function $\varphi=const$ is not very realistic.
%new text ends

The main part of this article is organized as follows. In Section 2, a description of the GSA is presented, and theoretical statements regarding its convergence are given. The modification of the algorithm that makes it applicable for solving a series of problems simultaneously is described in Section 3. A uniform convergence of the modified algorithm to solving a series of problems is substantiated in Section 4. Section 5 presents the results of the numerical experiments; a comparison is made of various parallelization methods between the processors when solving a series of multidimensional multiextremal test problems. Section 6 concludes the paper.

\section{Global search algorithm} \label{sec:1}

Let us consider a global optimization problem 
\begin{eqnarray}\label{stand_problem}
& \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
& D=\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1\leq i \leq N\right\}, \nonumber
\end{eqnarray}
with a Lipschitzian objective function. The issues related to reducing more general problem statements to standard statement (\ref{stand_problem}) were considered in \cite{Strongin2009}.

Using a continuous unambiguous mapping $y(x)$ of the interval $[0,1]$ of the real axis onto hypercube $D$ from (\ref{stand_problem}) (\textit{Peano space-filling curve}) one can reduce the multidimensional problem (\ref{stand_problem}) to a one-dimensional problem
\begin{equation}\label{u_problem}
\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}}.
\end{equation}

As follows from the properties of the space-filling curve \cite{Strongin2000}, the reduced one-dimensional function $\varphi(y(x))$ satisfies a uniform H{\"o}lder condition
\[
\left|\varphi(y(x'))-\varphi(y(x''))\right|\leq H\left|x'-x''\right|^{1/N},
\]
where the H{\"o}lder constant $H$ is related to the Lipschitz constant $L$ by the relationship 
\begin{equation}\label{L_and_H}
H=2L\sqrt{N+3}.
\end{equation}
Problems of numerical construction of Peano-type space filling curves and the corresponding theory are considered in detail in \cite{Strongin2000,Sergeyev2013}. Here we will note that a numerically constructed curve (\textit{evolvent}) is $2^{-m}$ accurate approximation of the theoretical Peano curve in $L_\infty$ metric, where $m$ is an evolvent construction parameter.

The algorithm considered here generates a series of preimages  $x^k\in[0,1], k=0,1,2,...$, in which first the images $y^k=y(x^k)$ and then the values of the objective function $z^k=\varphi(y^k), k=0, 1, 2, ...$,  are computed. Let us call the process of computing a function value (including the construction of the image $y^k=y(x^k)$) a \textit{trial}, and the pair $(x^k, z^k)$ the result of the trial. The set of pairs
\[
\omega_k=\left\{(x^i,z^i):0\leq i \leq k\right\},
\]
corresponding to the first $k$ terms of the sequence of preimages ${x^k}$ is the \textit{search information} accumulated by the algorithm within $k$ steps.

According to the global search algorithm, the first two trials are executed at the points 
\[
y^0=y(0), y^1=y(1).
\]
The choice of the point $y^{k+1},k\geq 1,$  for the next $(k+1)$-th trial is defined by the following rules.

\begin{enumerate}
	\item 
	Renumber the inverse images of all the points from the trials already performed  
\begin{equation}\label{y_i}
y^0=y(x^0), y^1=y(x^1),...,y^k=y(x^k)
\end{equation}
by subscripts in the increasing order of their coordinates, i.e.
\begin{equation}\label{x_i}
0=x_0<x_1<\dots <x_k=1,
\end{equation}
and associate these with the values $z_i=\varphi(y(x_i)), 0\leq i \leq k,$  computed at these points.
\item
Compute the maximum absolute value of the first divided differences
\begin{equation}\label{mu}
\mu = \max_{1 \leq i \leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
\end{equation}
where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$. If $\mu = 0$, set $\mu = 1$.
\item
For each interval $(x_{i-1}, x_i), \; 1\leq i \leq k,$  calculate the value
\begin{equation}\label{R}
R(i)=r\mu\Delta_i+\frac{(z_i-z_{i-1})^2}{r\mu\Delta_i}-2(z_i+z_{i-1})
\end{equation}
called the \textit{characteristic} of the interval; the real number $r>1$ being the input parameter of the algorithm.
\item 
Select the interval $(x_{t-1},x_t)$ corresponding to the maximum characteristic
\begin{equation}\label{MaxR}
R(t)= \max_{1 \leq i \leq k}R(i).
\end{equation}
If there are several such intervals, choose the minimum number satisfying condition (\ref{MaxR}) as $t$.
\item
Assume
\begin{equation}\label{xk1}
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N
\end{equation}
as the inverse image for the next trial point, i.e. the next trial is to be carried out at the point $y^{k+1}=y(x^{k+1})$ from $D$.
\end{enumerate}

Rules (\ref{y_i})--(\ref{xk1}) describe the sequence of decision functions 
\[
x^{k+1}=G_k^r(x^0,...,x^k,z^0,...,z^k), \; k=1,2,...,
\]
generating the sequence of inverse images $x^k\in[0,1]$, $x^{k+1} \notin \left\{x^0,...,x^k\right\}$, and also the sequence of trial points $y^k\in D$, $y^{k+1} \notin \left\{y^0,...,y^k\right\}$. This infinite sequence of trials may be terminated by meeting the stopping condition
\[
\Delta_t < \epsilon,
\]
where $t$ is from (\ref{MaxR}), and $\epsilon>0$ is the predefined accuracy. We will use the value
\[
\varphi_k^\ast=\min_{0\leq i \leq k}\varphi(y^i)
\]
and the point 
\[
y_k^\ast=\arg \min_{0\leq i \leq k}\varphi(y^i)
\]
as the estimate of the optimum corresponding to the $k$-th step of the algorithm.

The conditions for convergence of the algorithm described above are formulated in the form of the following theorem from \cite{Strongin2000}.

\textbf{Theorem 1} (sufficient convergence conditions)

Let the point $\overline{y}$ be a limit point of the sequence $\left\{y^k\right\}$ generated by the rules of GSA while minimizing a Lipschitzian function $\varphi(y)$, $y \in D$, with the constant $L$. Then:
\begin{enumerate}
	\item 
	If side by side with $\overline{y}$ there exists another limit point $y'$ of the sequence $\left\{y^k\right\}$, then $\varphi (\overline{y}) = \varphi (y')$.
	\item
	For any  $z^k=\varphi(y^k) \geq \varphi(\overline{y})$.
	\item
	If at the some step of the search process the value $\mu$ from (\ref{mu}) satisfies the condition
\begin{equation}\label{217}
r \mu > 2^{3-1/N}L\sqrt{N+3},
\end{equation}
then $\overline{y}$ is a global minimizer of the function $\varphi (y)$ over $D$ and any global minimizer $y^\ast$ from (\ref{stand_problem}) is also a limit point of the sequence $\left\{y^k\right\}$.
\end{enumerate}

\textbf{Remark 1}. It follows from relationship (\ref{L_and_H}) between the Lipschitz and H{\"o}lder constants, as well as from condition (\ref{217}) of the Theorem, that for the convergence of the algorithm, parameter $r$ should satisfy the condition
\begin{equation}\label{218}
r > 2^{2-1/N}.
\end{equation}

\textbf{Remark 2}. Substantiation of the theorem is obtained by reducing to the contradiction with condition (\ref{MaxR}), i.e. by deriving the consequence that the subsequent trial will be carried out within an interval without the maximum characteristic.

The algorithm considered above is very flexible and allows an efficient parallelization. Let us assume a multiprocessor system with $p$ cores/processors to be at our disposal. The computational complexity of the objective function is one of the basic assumptions (see Introduction). The computations related to applying the algorithm's decision rules will play a role with negligible overhead costs. In this regard, the execution of $p$ search trials in parallel on different cores/processors within one iteration of the method would be the most efficient parallelization scheme. Thus the obtained parallel algorithm will consist of the following.

The first $p$ trials are executed at the points $x^0 = 0$, $x^1 = 1$ and at the arbitrary internal points $x^2, ..., x^{p-1}$ of the interval $(0,1)$. Let us assume $n \geq 1$  iterations of the method to be completed, in the course of which the trials in $k=k(n)$ points $x^i, 0 \leq i \leq k,$ have been executed. Then, the points $x^{k+1},...,x^{k+p}$  of the search trials for the next $(n+1)$-th iteration are determined according to the following rules.

Steps 1--3 of the parallel algorithm repeat steps 1--3 of GSA completely.

4. Arrange characteristics  $R(i), 1 \leq i \leq k$, in decreasing order 
\begin{equation}\label{Rdec}
R(t_1)\geq R(t_2)\geq \dots \geq R(t_{k})
\end{equation}
and select $p$ largest characteristics with interval numbers $t_j, 1\leq j \leq p$.

5. Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), 1\leq j\leq p$, computed according to the formula
\[
x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - \mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N.
\]

The algorithm stops if condition $\Delta_{t_j}<\epsilon$ is satisfied for at least one index $t_j, 1 \leq j \leq p$ ; here $\epsilon>0$ is the predefined accuracy.

The decision rules of the Parallel Global Search Algorithm (PGSA) allow the following interpretation (see \cite{Strongin2000}), according to which series (\ref{Rdec}) arranges the intervals $(x_{t-1},x_t), 1 \leq t \leq k,$ with respect to the decreasing probability of the localization of the global minimizer of the function $\varphi(y(x))$ in these intervals. Therefore, (as part of this interpretation) one can expect the next trial points to be chosen within the intervals with the largest probabilities at the current search step.

The theorem on the convergence of PGSA is a generalization of theorem 1, of which is also presented in \cite{Strongin2000}. In particular, it has been proven that the introduction of the parallelism described above does not produce limit points different from those in a purely sequential scheme. The results of the investigation of sequential and parallel algorithms (as well as their various modification) are presented in \cite{Sergeyev1994,Grishagin1997}.

As an example of the work of GSA and PGSA, let us consider the minimization of a two-dimensional function obtained using a GKLS generator \cite{Gaviano2003}. In Fig.~\ref{example} (a) and (b), the contour plots of the objective function are presented. Fig.~\ref{example} (a) also shows the points of 261 trials (261 iterations) performed by the sequential version of the method until the required accuracy $\epsilon=10^{-2}$ was achieved. In Fig.~\ref{example} (b) the points of 304 trials (76 iterations) performed by the parallel algorithm using 4 cores, i.e. at $p=4$, are shown. The points of the trials performed by the parallel algorithm which differ from the trial points of the sequential algorithm are marked by the '+' sign.

\begin{figure}
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig2a.JPG} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig2b.JPG} \\ (b)}
\end{minipage}
\caption{Solving a two-dimensional problem using (a) GSA and (b) PGSA}
\label{example}
\end{figure}

\section{An algorithm for solving a series of optimization problems} \label{sec:2}

Now, let us assume that we need to solve a series of $q$ problems 
\[
\min\left\{\varphi_1(y), y\in D \right\}, \min\left\{\varphi_2(y), y\in D\right\},..., \min\left\{\varphi_q(y), y\in D\right\}.
\]

It is necessary to introduce some modifications to the algorithm's computational scheme which would allow obtaining the estimates of optima in all the problems simultaneously. It is also necessary to provide the properties of the trial sequences generated during the minimization of different functions that do not depend on a particular form of the function.

The fulfillment of the inequality  
\begin{equation}\label{4phi}
R(j(l))>-4\varphi(y(x^\ast)),
\end{equation}
where $R(j(l))$ is the characteristic of the interval $(x_{j-1},x_j), \; j=j(l),$ containing the global minimizer $x^\ast$ of the function $\varphi(y(x))$ during the $l$-th iteration, is one of key properties used in \cite{Strongin2000} in the proof for the algorithm's convergence.

As seen from the inequality (\ref{4phi}), the value of the characteristic of the interval $(x_{j-1},x_j), \; j=j(l),$ depends on the value of the function. If the algorithm is applied to minimize different functions, the values of the characteristics of intervals containing the global minimizer will be different as well. Hence in order to use the algorithm to minimize different functions, the algorithm's decision rules should be modified so that the value of the characteristic for the best interval does not depend on a particular form of the function. Let us consider such modifications.

1. If all characteristics $R(i)$ from (\ref{R}) at a given $k$ are divided by a quantity $r\mu$, where $\mu$ is from (\ref{mu}), i.e. if the characteristics are computed according to the formula
\[
R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{(r\mu)^2\Delta_i}-\frac{2(z_i+z_{i-1})}{r\mu},
\]
the relationship of inequality between the characteristics would not change.

2. If a quantity $4\varphi_\ast^k/r\mu$  is added to all characteristics at a given $k$, the relationship of inequality between these would not change either.
Thus, the algorithms with characteristics (\ref{R}) and with characteristics
\begin{equation}\label{NewR}
R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{(r\mu)^2\Delta_i}-\frac{2(z_i+z_{i-1}-2\varphi_\ast^k)}{r\mu}
\end{equation}
will generate the same sequences $\left\{x^k\right\}$. Consequently, the theorem of convergence is true for an algorithm with the characteristics (\ref{NewR}) since the normalization of the characteristics at each step does not change the relationship of their equality and superiority. Then, the inequality (\ref{4phi}) transforms into the inequality
\begin{equation}\label{zero}
R(j(l))>0,
\end{equation}
i.e. the lower bound for the best interval characteristic does not depend on the objective function now.

\section{Conditions of uniform convergence for the algorithm}

Let us consider the problems in the form  (\ref{stand_problem}) for a set of $q$ lipschitzian functions $\varphi_1(y), \varphi_2(y),...,\varphi_q(y)$ defined in $D$. Let us select (arbitrarily) two of these problems:
\begin{eqnarray} \label{two_problems}
\min\left\{\varphi(y), y\in D\right\}, \\ \nonumber
\min\left\{\psi(y), y\in D\right\}. \nonumber
\end{eqnarray}
Let us denote the characteristics (\ref{NewR}) for the first problem as $R_\varphi(i)$ and for the second problem as $R_\psi(j)$. Given that, 
\begin{eqnarray} \nonumber
R_\varphi(t_\varphi)= \max_{1 \leq i \leq k}R_\varphi(i), \\ \nonumber
R_\psi(t_\psi)= \max_{1 \leq j \leq s}R_\psi(j), \nonumber
\end{eqnarray}
where $k$ corresponds to the number of trials in the first problem and $s$ corresponds to the number of trials in the second one. The sequence of trials $\left\{v^k\right\}$ corresponds to the first problem and the sequence of trials $\left\{u^s\right\}$ corresponds to the second problem. The values $z^k=\varphi (y(v^k))$ correspond to the trial points $v^k\in [0,1]$ and the values $w^s=\psi(y(u^s))$ correspond to the trial points $u^s \in [0,1]$.

Let us consider the case where these two problems are solved simultaneously, and the selection of the interval $(v_{t_\varphi-1},v_{t_\varphi})$ or $(u_{t_\psi-1},u_{t_\psi})$ to execute the trials is determined by the condition
\begin{equation}\label{NewMaxR}
R(t)=\max\{R_\varphi(t_\varphi),R_\psi(t_\psi)\}.
\end{equation}

\textbf{Theorem 2} (on the convergence conditions)

Let the conditions for theorem 1 be satisfied for each problem (\ref{two_problems}) separately. 
Then solving both problems by the algorithm defining the interval for each next trial according to rule (\ref{NewMaxR}) for characteristics (\ref{NewR}) will generate two infinite sequences $\left\{v^k\right\}$ and $\left\{u^s\right\}$, for the limit points of which theorem 1 is also true.

\textbf{Proof}. Let $l=k+s,l=0,1,2,...,$ and the condition $v^0=u^0=0$, $v^1=u^1=1$ be satisfied. Then, the sequence of trials (i.e. $\left\{y(v^k)\right\}$) will be infinite at least for one of the functions (let it be $\varphi(y)$), and theorem 1 will be true for this function. We show that the algorithm generates an infinite sequence $\left\{y(u^s)\right\}$ for the second function $\psi(y)$ as well, for which, consequently, theorem 1 is  also true.

Let the limit point $\overline{v}\in[v_{i-1},v_i]$ where $i=i(k)$ and the parameter of algorithm $r$ satisfies the condition $r>2^{2-1/N}$, according to (\ref{218}). From (\ref{mu})
\[
\mu_\varphi^2 \geq \left(\frac{|z_i-z_{i-1}|}{\Delta_i}\right)^2
\]
and, consequently,
\[
\frac{(z_i-z_{i-1})^2}{r^2\mu_\varphi^2\Delta_i}=\frac{(z_i-z_{i-1})^2\Delta_i}{(r\mu_\varphi\Delta_i)^2}\leq\frac{\Delta_i}{r^2}.
\]
Hence, according to (\ref{NewR})
\[
R_\varphi(i) \leq \Delta_i\left(1+\frac{1}{r^2}\right)-\frac{2(z_i+z_{i-1}-2\varphi_\ast^k)}{r\mu_\varphi}.
\]
Because $\overline{v}$ is the limit point of the sequence of preimages $\left\{v^k\right\}$ and $\varphi_\ast^k \geq \varphi(y(\overline{v}))$ at $k\rightarrow\infty$
\begin{eqnarray} \nonumber
\Delta_i\rightarrow 0, \\ \nonumber
z_i+z_{i-1}-2\varphi_\ast^k\rightarrow 0. \nonumber
\end{eqnarray}
Therefore, for any small $\delta>0$ there exists a large value of $k$ such that 
\begin{equation}\label{Rleqdelta}
R_\varphi(i)<\delta.
\end{equation}

Let $\psi_\ast^s=w_{j-1}$, i.e. the current minimum value of function $\psi(y)$ is achieved at the left point of the interval $(w_{j-1},w_j)$. Then 
\[
\frac{w_j+w_{j-1}-2\psi_\ast^s}{r\mu_\psi}=\frac{w_j-w_{j-1}}{r\mu_\psi} = \frac{\Delta_j(w_j-w_{j-1})}{r\mu_\psi \Delta_j}\leq \frac{\Delta_j}{r}
\]
and, taking into account that $r>2^{2-1/N}>2$,
\[
R_\psi(j)=\Delta_j+\frac{(w_j-w_{j-1})^2}{r^2\mu_\psi^2\Delta_j}-\frac{2(w_j+w_{j-1}-2\psi_\ast^s)}{r\mu_\psi} \geq \Delta_j\left(1-\frac{2}{r}\right) > 0.
\]
Hence, according to (\ref{Rleqdelta}), the inequality 
\[
R_\psi(j) > R_\varphi(i).
\]
will be true at large enough values of $k$, and the scheduled iteration falls into the interval $(u_{j-1},u_j)$, i.e. the sequence $\left\{u^s\right\}$ will be infinite as well.

\textbf{Theorem 3} (on the uniform convergence)

Let the conditions of theorem 2 be satisfied. Then the convergence to the limit points of the sequences $\left\{v^k\right\}$ and $\left\{u^s\right\}$ is uniform,
%new text begins
i.e. for any step $l=k+s$  the minimum distance between the points of the sequence $\left\{v^k\right\}$ and its limit point does not exceed a similar value for the sequence $\left\{u^s\right\}$. 
%new text ends

\textbf{Proof}. Let $v^\ast$ be the preimage of the global optimizer of the function $\varphi(y)$, and $u^\ast$ be the preimage of the global optimizer of the function $\psi(y)$, given that 
\[
v^\ast \in \left[v_{i-1},v_i\right], \; u^\ast \in \left[u_{j-1},u_j\right],
\]
and let 
\begin{equation}\label{RpsileqRphi}
R_\psi(j) > R_\varphi(i)
\end{equation}
at some step $l=k+s$. Consequently, the next trial will fall into the interval $(u_{j-1},u_j)$ before the interval $(v_{i-1},v_i)$. At large enough values of $s$, it follows from the convergence of $\left\{u^s\right\}$ to $u^\ast$ that $\psi_\ast^s=w_{j-1}$, whence 
\[
w_j+w_{j-1}-2\psi_\ast^s=w_j+w_{j-1}-2w_{j-1}=w_j-w_{j-1}
\]
and
\begin{equation}\label{Rpsij}
R_\psi(j) =\Delta_j+\Delta_j\left(\frac{w_j-w_{j-1}}{r\mu_\psi\Delta_j}\right)^2-2\Delta_j\frac{w_j-w_{j-1}}{r\mu_\psi\Delta_j}=\Delta_j\left(1-\frac{w_j-w_{j-1}}{r\mu_\psi\Delta_j}\right)^2,
\end{equation}
where $\Delta_j=(u_j-u_{j-1})^{1/N}$. Analogously, if $\varphi_\ast^k=z_{i-1}$, then
\begin{equation}\label{Rphii}
R_\varphi(i) = \Delta_i\left(1-\frac{z_i-z_{i-1}}{r\mu_\varphi\Delta_i}\right)^2,
\end{equation}
where $\Delta_i=(v_i-v_{i-1})^{1/N}$. Now, it follows from (\ref{RpsileqRphi})--(\ref{Rphii}) that
\[
\Delta_j\left(1-\frac{w_j-w_{j-1}}{r\mu_\psi\Delta_j}\right)^2 > \Delta_i\left(1-\frac{z_i-z_{i-1}}{r\mu_\varphi\Delta_i}\right)^2
\]
and
\[
\frac{\Delta_i}{\Delta_j} <  \frac{\left(1-\frac{w_j-w_{j-1}}{r\mu_\psi\Delta_j}\right)^2}{\left(1-\frac{z_i-z_{i-1}}{r\mu_\varphi\Delta_i}\right)^2}.
\]
Because 
\[
0 \leq \left|\frac{w_j-w_{j-1}}{r\mu_\psi\Delta_j}\right| \leq \frac{1}{r}, \; 0 \leq \left|\frac{z_i-z_{i-1}}{r\mu_\varphi\Delta_i}\right| \leq \frac{1}{r},
\]
then
\[
\frac{\Delta_i}{\Delta_j} <  \frac{\left(1+\frac{1}{r}\right)^2}{\left(1-\frac{1}{r}\right)^2}=\left(\frac{r+1}{r-1}\right)^2.
\]

Thus, at the moment when the next trial falls into the interval $(u_{j-1},u_j)$, the length of the interval $(v_{i-1},v_i)$ cannot exceed the length of the interval $(u_{j-1},u_j)$ by more than $\left(\frac{r+1}{r-1}\right)^2$ times. The uniform convergence has been proven.

\textbf{Remark 3}. It follows from the properties of the evolvent (see \cite{Sergeyev2013}) that any two close points $x',x''\in (0,1)$ correspond to close images $y'=y(x')$, $y''=y(x'')$ in the hypercube $D$, and if
\[
\left|x'-x''\right|\leq 2^{-N(m+1)},
\]
where $m\geq 1$, then
\[
\max \left\{  \left|y'_i-y''_i\right|: 1\leq i \leq N \right\} \leq 2^{-m}.
\]
Thus, the uniform convergence of the images in the multidimensional space also follows from the uniform convergence of the preimages on the one-dimensional interval as well.

\textbf{Remark 4}. Since two problems (\ref{two_problems}) were selected from a set of problems being solved arbitrarily, theorem 2 and theorem 3 will also be true for any pair of problems from the set being solved. At the same time, theorem 2 and theorem 3 will also be true for $q>2$ problems 
\[
\min\left\{\varphi_1(y), y\in D \right\}, \; \min\left\{\varphi_2(y), y\in D\right\},..., \; \min\left\{\varphi_q(y), y\in D\right\}
\]
being solved simultaneously by induction.

\textbf{Remark 5}. The theorems on uniform convergence for the parallel algorithm are proved similarly, which allows applying the multiprocessor systems to solve a set of problems simultaneously.

\textbf{Remark 6}. The parallel algorithm will provide a uniform workload distribution for the cores/processors used. We assume that computing the objective function value will be a time-consuming operation, and it accounts for the majority of the computation costs as compared to the application of the parallel algorithm decision rules (the computing and arrangement of the characteristics, the information interchange between parallel processes, etc.). However, just this most computationally costly part is parallelized ($p$ trials are executed at each iteration where $p$ is the number of cores/processors employed). Therefore, the load on the computer system will be close to its full capacity.

\section{Results of experiments}

One of the well-known approaches to investigating and comparing multiextremal optimization algorithms is based on the application of these methods for solving a set of test problems selected randomly from some constructed class. In this case, each test problem can be considered as a particular realization of a random function defined by means of a special generator. The application of multiextremal optimization algorithms to sets of these functions allows the efficiency of each particular algorithm to be evaluated.

The numerical comparison of the algorithms has been carried out using the GKLS test problem generator \cite{Gaviano2003}. This generator allows multiextremal optimization problems to be generated with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.). Eight GKLS classes of differentiable test functions of the dimensions $N = 2,3,4,$ and $5$ have been used. For each dimension, both \textit{Hard} and \textit{Simple} classes have been considered. The difficulty of a class was increased either by decreasing the radius of the attraction region of the global minimizer, or by decreasing the distance from the global minimizer $y^\ast$ to the domain boundaries. The global minimizer $y^\ast$ was considered to be found, if the algorithm generated a trial point $y^k$ in the vicinity of the global minimum, i.e. $\left|y^k-y^\ast\right| < \delta \left\|b-a\right\|$, where $\delta = 0.01$, $a$ and $b$ are the boundaries of the search domain $D$. When using the Global Search Algorithm, the parameter $r$ was set to $5$, the evolvent construction parameter was fixed as $m = 10$. The maximum allowable number of iterations per problem was $K_{max} = 10^6$.

Each series of problems was solved using a scheme to distribute the problems between the parallel threads ($S1$) and using the proposed scheme to simultaneously solve all problems ($S2$). The number of employed parallel threads $p$ was varied from $1$ to $24$.

The average numbers of trials executed by the algorithm depending on the number of threads when solving \textit{Simple} and \textit{Hard} classes of problems are presented in Tables \ref{tab:1} and \ref{tab:2}, respectively. The results demonstrate that both schemes of problem distribution between threads require an almost equal number of trials. The workload of each thread is also the same. In particular, there is an exact match between the number of trials for the sequential runs (at $p = 1$).

\begin{table}
	\caption{Average number of iterations for the \textit{Simple} class}
	\label{tab:1}
	\center
	\begin{tabular}{cccccccccccc}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{c}{ $N=2$ } & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12}  \noalign{\smallskip}
		 & $S1$ & $S2$ & & $S1$ & $S2$ & & $S1$ & $S2$ & & $S1$ & $S2$  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
1	&	408	&	408	&	&	2502	&	2502	&	&	28254	&	28254	&	&	87261	&	87261	\\
2	&	408	&	415	&	&	2502	&	2496	&	&	28254	&	27206	&	&	87261	&	85994	\\
4	&	408	&	405	&	&	2502	&	2430	&	&	28254	&	30121	&	&	87261	&	78064	\\
8	&	408	&	402	&	&	2502	&	2505	&	&	28254	&	28980	&	&	87261	&	84337	\\
16	&	408	&	415	&	&	2502	&	2501	&	&	28254	&	28206	&	&	87261	&	82564	\\
24	&	408	&	400	&	&	2502	&	2416	&	&	28254	&	27868	&	&	87261	&	86252	\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Average number of iterations for the \textit{Hard} class}
	\label{tab:2}
	\center
	\begin{tabular}{cccccccccccc}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{c}{ $N=2$ } & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12}  \noalign{\smallskip}
		 & $S1$ & $S2$ & & $S1$ & $S2$ & & $S1$ & $S2$ & & $S1$ & $S2$  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
1	&	806	&	806	&	&	3873	&	3873	&	&	60972	&	60972	&	&	162475	&	162475	\\
2	&	806	&	807	&	&	3873	&	4104	&	&	60972	&	58382	&	&	162475	&	167728	\\
4	&	806	&	823	&	&	3873	&	4081	&	&	60972	&	57285	&	&	162475	&	172522	\\
8	&	806	&	881	&	&	3873	&	4153	&	&	60972	&	55353	&	&	162475	&	173725	\\
16	&	806	&	854	&	&	3873	&	4165	&	&	60972	&	60158	&	&	162475	&	178130	\\
24	&	806	&	887	&	&	3873	&	4270	&	&	60972	&	55939	&	&	162475	&	182821	\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

In order to demonstrate the presence of the uniform convergence of the proposed algorithm for solving a series of problems, let us plot the functions $D_{av}(K)$ and $D_{max}(K)$ featuring the average and maximum deviation of the current approximation from the exact solution over all problems, depending on the total number of trials. 
Figs.~\ref{fig_last1},~\ref{fig_last2},~\ref{fig_last3} and \ref{fig_last4} show these functions plotted while solving five-dimensional \textit{Simple} class problems using a sequential algorithm and a parallel algorithm.
The number of trials is plotted on the abscissa (in millions) against the deviation of the current approximation from the exact solution of the problem on the ordinate. The red curve shows the first scheme of problem distributions while the green curve represents the second scheme.

\begin{figure}
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig3a.jpg} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig3b.jpg} \\ (b)}
\end{minipage}
\caption{Average (a) and maximum (b) deviation of the current approximation from the exact solution for the sequential algorithm.}
\label{fig_last1}
\end{figure}

\begin{figure}
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig4a.jpg} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig4b.jpg} \\ (b)}
\end{minipage}
\caption{Average (a) and maximum (b) deviation of the current approximation from the exact solution for the parallel algorithm; the number of threads $p = 2$.}
\label{fig_last2}
\end{figure}

\begin{figure}
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig5a.jpg} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig5b.jpg} \\ (b)}
\end{minipage}
\caption{Average (a) and maximum (b) deviation of the current approximation from the exact solution for the parallel algorithm; the number of threads $p = 8$.}
\label{fig_last3}
\end{figure}

\begin{figure}
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig6a.jpg} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig6b.jpg} \\ (b)}
\end{minipage}
\caption{Average (a) and maximum (b) deviation of the current approximation from the exact solution for the parallel algorithm; the number of threads $p = 24$.}
\label{fig_last4}
\end{figure}

The plots clearly demonstrate the uniform convergence to the solutions of the problems of the series when using Scheme 2. The arrangement of curves in Figs.~\ref{fig_last1},~\ref{fig_last2},~\ref{fig_last3} and \ref{fig_last4} demonstrates that when simultaneously solving the problems, one can obtain a good approximation to the solution in all problems within $100 000$ trials (on average as well as in the worst case). In both cases the scheme of problem distribution between the independent threads is inferior to the scheme for simultaneous solving of all problems.

%new text begins

Previous experiments were carried out with the objective functions of the same structure. To demonstrate the uniform convergence for objective functions of a different structure let us consider two mixed series of two-dimensional problems: 50 functions from GKLS \textit{Simple} or \textit{Hard} classes and 50 functions of the form
\begin{eqnarray} \nonumber \label{vagris}
\varphi(y)= -&\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2+\right. \\
&\left.\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},\\ \nonumber
\end{eqnarray}
where
\begin{eqnarray} \nonumber
& y=(y_1,y_2)\in R^2, 0 \leq y_1,y_2 \leq 1, \\ \nonumber
& g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2),  \\ \nonumber
& h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2), \nonumber 
\end{eqnarray}
and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$. The problems with odd numbers are from GKLS classes and the problems with even numbers are based on functions (\ref{vagris}).

Each series of problems was solved using a scheme for distributing the problems between the parallel threads ($S1$) and using the scheme for solving all problems simultaneously ($S2$). The number of employed parallel threads $p$ was varied from 1 to 16. The average numbers of trials executed by the algorithm depending on the number of threads when solving mixed classes of problems are presented in Table \ref{tab:3}.

\begin{table}
	\caption{Average number of iterations for the mixed series of problems}
	\label{tab:3}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{c}{ GKLS \textit{Simple} and (\ref{vagris}) } & & \multicolumn{2}{c}{GKLS \textit{Hard} and (\ref{vagris})} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & $S1$ & $S2$ & & $S1$ & $S2$ \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
1	&	234	&	234	&	&	472	&	472	\\
2	&	234	&	240	&	&	472	&	476	\\
4	&	234	&	248	&	&	472	&	469	\\
8	&	234	&	249	&	&	472	&	482	\\
16	&	234	&	251	&	&	472	&	479	\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

Figs. \ref{fig_last5}, \ref{fig_last6}, \ref{fig_last7} show the functions $D_{av}(K)$ and $D_{max}(K)$ plotted when solving mixed problem series using the sequential algorithm and the parallel algorithm. The number of trials is plotted on the abscissa (in thousands) against the deviation of the current approximation from the exact solution of the problem on the ordinate. The red curve shows the first scheme of problem distributions while the green curve represents the second scheme.

\begin{figure}
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig7a.jpg} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig7b.jpg} \\ (b)}
\end{minipage}
\caption{Average (a) and maximum (b) deviation of the current approximation from the exact solution for the sequential algorithm when solving mixed series of problems.}
\label{fig_last5}
\end{figure}

\begin{figure}
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig8a.jpg} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig8b.jpg} \\ (b)}
\end{minipage}
\caption{Average (a) and maximum (b) deviation of the current approximation from the exact solution for the parallel algorithm when solving mixed series of problems; the number of threads $p = 8$.}
\label{fig_last6}
\end{figure}

\begin{figure}
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig9a.jpg} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig9b.jpg} \\ (b)}
\end{minipage}
\caption{Average (a) and maximum (b) deviation of the current approximation from the exact solution for the parallel algorithm when solving mixed series of problems; the number of threads $p = 16$.}
\label{fig_last7}
\end{figure}

As in the case of objective functions of the same structure, the results show the uniform convergence when using the scheme for solving all problems simultaneously.

%new text ends

The computational experiments were carried out on one of the nodes of a high-performance cluster at Lobachevsky State University of Nizhny Novgorod. The cluster node included 2 Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The CPU had 8 cores, i.e. a total of 16 physical cores and 16 virtual cores in the hyper-threading mode were available in the node. MS Visual Studio 15 was used to implement the algorithm.

\section{Conclusions }

In this paper, the results obtained when investigating the issues in solving a series of optimization problems are presented. A series of problems may arise when solving a more complex type of problem, for example, a multiple-criteria optimization problem or a larger dimensional one. A modification of the Global Search Algorithm has been proposed to allow several problems to be solved at once and, correspondingly, simultaneously obtain estimates for the solutions to these problems. The theorem regarding the uniform convergence of the novel algorithm has been proven. The computational experiments on a series of several hundred test problems of various dimensions have been carried out. The experiments have clearly demonstrated the uniform convergence.

%Further generalization of the proposed approach onto a wider class of the parallel characteristical algorithms of global optimization \cite{Grishagin1997} seems promising. It would expand its sphere of application essentially since many well-known global optimization algorithms can be represented in the characteristical form.

% Non-BibTeX users please use
\begin{thebibliography}{}

\bibitem{Strongin2000}
Strongin, R.G., Sergeyev, Ya.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000)

\bibitem{Sergeyev2013}
Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting Space-Filling Curves. Springer (2013)

\bibitem{Pinter1996}
Pinter, J. D.: Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications). Kluwer Academic Publishers, Dordrecht (1996)

\bibitem{Jones2001}
Jones, D. R.: The direct global optimization algorithm. In: Floudas, C. A., Pardalos, P. M. (eds.) The Encyclopedia of Optimization, Second Edition. pp. 725--735. Springer (2009)

\bibitem{Evtushenko2013}
Evtushenko, Y. G., Posypkin, M. A.: A deterministic approach to global box-constrained optimization. Optim. Lett. 7(4), 819--829 (2013)

\bibitem{Zilinskas2008}
Zilinskas, J.: Branch and bound with simplicial partitions for global optimization. Mathematical Modelling and Analysis 13(1), 145--159 (2008)

\bibitem{Zilinskas2010}
Paulavicius, R., Zilinskas, J., Grothey, A.: Investigation of selection strategies in branch and bound algorithm with simplicial partitions and combination of Lipschitz bounds. Optim. Lett. 4(2), 173–--83 (2010)

\bibitem{Sergeyev2000}
Kvasov, D.E., Pizzuti, C., Sergeyev, Y.D.: Local tuning and partition strategies for diagonal GO methods. Numer. Math. 94(1), 93--106 (2003)

\bibitem{Sergeyev2006}
Sergeyev, Ya.D., Kvasov, D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim. 16(3), 910--937 (2006)

\bibitem{Paulavicius2014}
Paulavicius, R., Sergeyev, Y.D., Kvasov, D.E., Zilinskas, J. Globally-biased \textsc{Disimpl} algorithm for expensive global optimization. J. Glob. Optim. 59(2), 545--567 (2014)

\bibitem{Gergel2005}
Gergel, V.P., Strongin, R.G.: Parallel computing for globally optimal decision making on cluster systems. Future Gener. Comput. Syst., 21(5), 673--678 (2005)

\bibitem{Evtushenko2009}
Evtushenko, Yu.G., Malkova, V.U., Stanevichyus, A.A.: Parallel global optimization of functions of several variables. Comput. Math. Math. Phys. 49(2), 246--260 (2009)

\bibitem{He2008}
He, J., Verstak, A., Watson, L.T., Sosonkina, M.: Design and implementation of a massively parallel version of DIRECT. Comput. Optim. Appl. 40(2), 217--245 (2008)

\bibitem{Paulavicius2011}
Paulavicius, R., Zilinskas, J., Grothey, A.: Parallel branch and bound for global optimization with combination of Lipschitz bounds. Optim. Methods Softw. 26(3), 487--498, (2011)

\bibitem{Ehrgott2005}
Ehrgott, M.: Multicriteria Optimization. Springer-Verlag, Berlin Heidelberg (2005)

\bibitem{Sergeyev1994}
Sergeyev, Y. D., Grishagin, V. A. Sequential and parallel algorithms for global optimization. Optim. Methods Softw. 3, 111--124 (1994)

\bibitem{Grishagin1997}
Grishagin, V.A., Sergeyev, Ya.D., Strongin, R.G.: Parallel characteristical algorithms for solving problems of global optimization. J. Glob. Optim. 10(2), 185--206 (1997)

\bibitem{Sergeyev2001}
Sergeyev, Y.D., Famularo D., Pugliese P. Index Branch-and-Bound Algorithm for Lipschitz univariate global optimization with multiextremal constraints. J. Glob. Optim. 21(3), 317--341 (2001)

\bibitem{Barkalov2002}
Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)

\bibitem{Strongin2003}
Strongin, R.G., Sergeyev, Ya.D.: Global optimization: fractal approach and non-redundant parallelism. J. Glob. Optim. 27(1), 25--50 (2003)

\bibitem{Barkalov2014}
Barkalov, K.A., Gergel, V.P.: Multilevel scheme of dimensionality reduction for parallel global search algorithms. Proceedings of the 1st International Conference on Engineering and Applied Sciences Optimization -- OPT-i 2014, 2111--2124 (2014)

\bibitem{Barkalov2016}
Barkalov, K., Gergel, V.: Parallel global optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016)

\bibitem{Zilinskas2013}
Zilinskas, A.: On the worst-case optimal multi-objective global optimization. Optim. Lett. 7, 1921--1928 (2013)

\bibitem{Strongin2009}
Strongin, R.G.: Global Optimization Using Space Filling. In: Floudas, C. A., Pardalos, P. M. (eds.) The Encyclopedia of Optimization, Second Edition. pp. 1418-1423. Springer (2009)

\bibitem{Gaviano2003}
Gaviano, M., Lera, D., Kvasov, D.E., Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Trans. Math. Software 29, 469--480 (2003)

%\bibitem{RefSergeyev2001} Molinaro, A., Pizzuti, C., and Sergeyev, Ya.D.: Acceleration tools for diagonal information global optimization algorithms. Computational Optimization and Applications 18, 5--26 (2001)

%\bibitem{RefGergel1996} Gergel, V.P.: A method of using derivatives in the minimization of multiextremum functions. Computational Mathematics and Mathematical Physics 36(6), 729--742 (1996)

%\bibitem{RefGergel1997} Gergel, V.P.: A global optimization algorithm for multivariate functions with lipschitzian first derivatives. J. Glob. Optim.10(3), 257--281 (1997)

%\bibitem{RefSergeyev1994} Sergeyev, Ya.D., Grishagin, V.A.: A parallel method for finding the global minimum of univariate functions. J. Optim. Theory Appl. 80(3), 513--536 (1994)

%\bibitem{RefSergeyevGrishagin} Sergeyev, Ya.D., Grishagin, V.A.:  Sequential and parallel global optimization algorithms. Optimization Methods and Software. 3, 111--124 (1994)

%\bibitem{RefGergel1999} Gergel, V.P., Sergeyev, Ya.D.: Sequential and parallel algorithms for global minimizing functions with Lipschitzian derivatives. Computers and Mathematics with Applications, 37(4-5), 163--179 (1999)

%\bibitem{RefBarkalov2013} Barkalov, K., Polovinkin, A., Meyerov, I., Sidorov, S., Zolotykh, N.: SVM regression parameters optimization using parallel global search algorithm // Lecture Notes in Computer Science. 7979, 154--166 (2013)

%\bibitem{RefJones} Jones, D.R., Perttunen, C.D., Stuckman, B.E.: Lipschitzian optimization without the lipschitz constant. J. Optim. Theory Appl. 79(1), 157--181 (1993)

%\bibitem{RefGablonsky} Gablonsky, J.M., Kelley, C.T.: A locally-biased form of the DIRECT algorithm. J. Glob. Optim., 21(1), 27--37 (2001)

\end{thebibliography}

\end{document}
% end of file template.tex

### 2017-08-22 2017/LION_Strongin/LION2017_BarkalovStrongin.tex
Added items remotely
D:\Common\Barkalov\Публикации\2017 LION\Тестовые задачи\LION_Strongin

\documentclass{llncs}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage[misc,geometry]{ifsym}

\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage{epstopdf}
%\usepackage{epsfig}

\begin{document}

\mainmatter

\title{Test Problems for Parallel Algorithms of Constrained Global Optimization }
\author{Konstantin Barkalov \Letter \and Roman Strongin \\
\email{ barkalov@vmk.unn.ru, strongin@unn.ru}}

\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia}

\maketitle

\begin{abstract}
This work considers the problem of building a class of test problems for global optimization algorithms. The authors present an approach to building multidimensional multiextremal problems, which can clearly demonstrate the nature of the best current approximation, regardless of the problem’s dimensionality. As part of this approach, the objective function and constraints arise in the process of solving an auxiliary approximation problem. The proposed generator allows the problem to be simplified or complicated, which results in changes to its dimensionality and changes in the feasible domain. The generator was tested by building and solving 100 problems using a parallel global optimization index algorithm. The algorithm's results are presented using different numbers of computing cores, which clearly demonstrate its acceleration and non-redundancy.

\keywords global optimization, multiextremal functions, non-convex constraints.

\end{abstract}

\section{Introduction}

One of the general approaches to studying and comparing multiextremal optimization algorithms is based on applying these methods to solve a set of test problems, selected at random from a certain specially constructed class. In this case, each test problem can be viewed as a random function created by a special generator. Using multiextremal optimization algorithms with large samples of such problems allows the operating characteristics of the methods to be evaluated (the likelihood of properly identifying the global optimizer within a given number of iterations), thus characterizing the efficiency of each particular algorithm.

The generator for one-dimensional problems was suggested by Hill \cite{Hill}. These test functions are typical for many engineering problems; they are particularly reminiscent of reduced stress functions in problems with multiple concentrated loads (see \cite{Toropov} for example). Another widely known class of one-dimensional test problems is produced using a generator developed by Shekel \cite{Shekel}.

A special GLOBALIZER software suite \cite{Globalizer} was developed to study various one-dimensional algorithms with random samples of functions produced by the Hill and Shekel generators. A comprehensive description of this system, its capabilities and example uses is provided in \cite{Strongin2000}. It should be noted that the Hill functions were successfully used in the design of a one-dimensional constrained problem generator with controlled measure of a feasible domain \cite{Barkalov2002}.

Another generator for random samples of two-dimensional test functions, successfully used in the studies by a number of authors, was developed and investigated in \cite{Grishagin1997}--\cite{Grishagin2016}. A generator for functions with arbitrary dimensionality was suggested in \cite{Gaviano}. It was used to study certain multidimensional algorithms as described in \cite{Kvasov2006}--\cite{SergeyevKvasov2015}.
%new text
%Известные коллекции тестовых задач для алгоритмов условной глобальной оптмизации были предложены в []). 
Well-known collections of test problems for constrained global optimization algorithms were proposed in \cite{Pardalos1990,Floudas1999}.

All of these generators produce the function to be optimized. In the case when the dimensionality is greater than two the optimization process itself cannot be clearly observed. In this regard, it is interesting to examine a different approach, initiated in \cite{Strongin2000}. In this approach, the objective function appears as a solution to a certain supporting approximation problem, which allows the nature of the best current estimate and the final result to be observed, regardless of the number of variables. Complicating the problem statement (including non-convex constraints) results in an increase of its dimensionality. In fact, the proposed generator produces an approximation problem to which the objective function is related.

\section{Problem statement}\label{Sec:2}

Let's consider the mathematical model of a charged particle moving through a magnetic field along the $u$ axis in the form\begin{equation}\label{Eq:1}
m\ddot{u}=-eu+F	
\end{equation}
where $m>0$ is the particle’s mass, $u(t)$ is the particle's current position at a moment of time \mbox{$t\geq0$}, $-eu$ -- is an attractive force affecting the particle, $F$ is an external force applied to the particle along $u$ axis (control action). It is assumed that control action $F$ is a function of time and is represented as
\[ 
F=m\sum^{n}_{i=1}A_i\sin(\omega_it+\varphi_i).
\]
Here $n>0$ is the dimensionality of the vectors $\omega$ and determines the number of frequencies in the control action.

Substituting  $\omega^2_0=e/m$ the equation (\ref{Eq:1}) is reduced to a known equation of forced oscillations
\begin{equation}\label{Eq:2}
\ddot{u}+\omega^2_0u=\sum^{n}_{i=1}A_i\sin(\omega_it+\varphi_i).	
\end{equation}
The problem is to find own frequency, control action and initial conditions such that:
\begin{enumerate}
\item
at $t\in[a,b]$ the particle would deviate from the position $q_0$ by no more than $\delta>0$;
\item
at $t=t_1, t_2, t_3$, the particle would deviate from positions $q_1, q_2, q_3$ respectively by no more than $\delta>0$;
\item
at $t=t_3$ the particle speed would be maximized.
\end{enumerate}
This problem statement can be interpreted as follows: the trajectory of particle movement $u(t)$ shall pass within a ''tube'', then through the three ''windows'', with maximum slope in the last of the ''windows''. The illustration in Fig.~\ref{fig11} shows a graph of the function $u(t)$ of the solution to problem  (\ref{Eq:2}), which passes through the ''tube'' and ''windows'' shown in the chart by dashed lines.

\begin{figure}
\begin{center}
  \includegraphics[width=0.75\textwidth]{fig1.jpg} 
  \caption{Problem solution $u(t)$} \label{fig11} 
\end{center}
\end{figure}

Thus, for a particle of a given mass $m>0$ it is necessary to determine its own frequency $\omega_0$, the amplitudes $A_i$, frequencies $\omega_i$ and phases $\varphi_i$ of the control action, as well as the initial conditions $u_0=u(0), \dot{u}_0=\dot{u}(0)$ for the equation (\ref{Eq:2}), such that conditions 1--3 are true.

Using the following notation for the equation (\ref{Eq:2}) solution 
\begin{equation}\label{Eq:3}
u(t,\omega,c)=\sum^{n}_{i=0}\left[c_{2i+1}\sin(\omega_it) + c_{2i+2}\cos(\omega_it)\right],	
\end{equation}
where  $c=\left(c_1,\ldots,c_{2n+2}\right)$, $\omega=\left(\omega_0,\omega_1,\ldots,\omega_n\right)$, we can represent the original problem as a constrained maximization problem with parameters  $c$ and $\omega$:
\begin{eqnarray} \label{Eq:4}
  & \left| \dot{u}(t_3,\omega,c) \right| \rightarrow \max \nonumber \\
  & \left|u(t_i,\omega,c)-q_i\right|\leq\delta, \; i=1,2,3, \\
	& \left|u(t,\omega,c)-q_0\right|\leq\delta, \; t\in[a,b]. \nonumber 
\end{eqnarray}

Solving the optimization problem (\ref{Eq:4}) and finding the vectors $\omega$, $c$ the solution to the original problem (\ref{Eq:2}) can be written in accordance with the following relationships: 
\begin{eqnarray} \label{Eq:5}
  & u_0=\sum^{n}_{i=0}c_{2i+2}, \; \; \dot{u}_0 = \sum^{n}_{i=0}c_{2i+1}\omega_i \nonumber \\
  & A_i=\left(\omega^2_0-\omega^2_i\right)\sqrt{c^2_{2i+1}+c^2_{2i+2}}\;, \; 1\leq i\leq n, \\
	& \varphi_i=\arcsin\frac{c_{2i+2}}{\sqrt{c^2_{2i+1}+c^2_{2i+2}}}\;, \; 1\leq i\leq n. \nonumber 
\end{eqnarray}

As follows from formula (\ref{Eq:3}), the parameters $c$ are included in the equation solution linearly, and the parameters $\omega$ -- non-linearly. Given the constraints (\ref{Eq:4}) this allows the problem to be reformulated and $c$ to be found without using a numerical optimization method.

Let's consider a set of points $(\tau_j,u_j), 0\leq j\leq m$, with the coordinates defined as follows:
\begin{eqnarray} \label{Eq:6}
  & \tau_j=a+jh, \; u_j=q_0, \; 0\leq j\leq m-3,\\
  & \tau_{m-2}=t_1, u_{m-2}=q_1, \nonumber \\
	& \tau_{m-1}=t_2, u_{m-1}=q_2, \nonumber \\
	& \tau_{m}=t_3, u_{m}=q_3, \nonumber
\end{eqnarray}
where $h=(b-a)/(m-3)$, i.e. the first $m-3$ points are located at equal distances in the center of the ''tube'', the other three align with the centers of the ''windows''.

The requirement is that the trajectory of particle $u(t)$ passes ''near'' the points  $(\tau_j,u_j), 0\leq j\leq m$. If the measure of deviation from the points is defined as the sum of the squared deviations
\[
\Delta(c,\omega)= \sum^{m}_{j=0}\left[u_j-u(\tau_j,\omega,c)\right]^2, 
\]
then the parameters $c$ can be found (given fixed values of $\omega$), by solving  the least squares problem
\begin{equation}\label{Eq:7}
c^\ast(\omega) = \arg \min\Delta(c,\omega).	
\end{equation}
According to the least squares method, the solution to problem (\ref{Eq:7}) can be obtained by solving a system of linear algebraic equations regarding the unknown $c$, which can be done, e.g., by Gaussian elimination.

It should also be considered that the components of frequency vector $\omega$ can be placed in ascending order, so as to avoid duplicate solutions corresponding to the vector $\omega$ with similar components in different order. In addition, it is natural to assume that frequencies in the control action must not just be ordered but differ by a certain positive value, as an actual physical device can only generate control signals with a certain precision. This assumption can be represented in the form of a requirement for the following inequalities to be true:
\begin{equation}\label{Eq:8}
\omega_{i-1}(1+\alpha)-\omega_i(1-\alpha)\leq 0, \; 1\leq i \leq n.	
\end{equation}
Here $\alpha\in(0,1)$ is a parameter reflecting the precision of signal generation by the control device.

Then the original problem can be reformulated as follows:
\begin{equation}\label{Eq:9}
\omega^\ast = \arg \max_{\omega \in \Omega} \left| \dot{u}(t_3,\omega,c^\ast(\omega)) \right|
\end{equation}
\[
\omega_{i-1}(1+\alpha)-\omega_i(1-\alpha)\leq 0, \; 1\leq i \leq n,
\]
\[
\left|u(t_i,\omega,c^\ast(\omega))-q_i\right|\leq\delta, \; i=1,2,3,
\]
\[
\max_{t\in[a,b]}{u(t,\omega,c^\ast(\omega))}-\min_{t\in[a,b]}{u(t,\omega,c^\ast(\omega))}\leq\delta,
\]
where $c^\ast(\omega)$ is determined from (\ref{Eq:7}), and the number of constraints will be dependent on the number of frequencies $n$ in the control action.

Fig.~\ref{fig11} shows trajectory $u(t)$, which corresponds to the solution to problem (\ref{Eq:9}) with parameters $a=1$, $b=10$, $t_1=13$, $t_2=16.65$, $t_3=18$, $q_0=q_3=0$, $q_1=7.65$, $q_2=-9.86$. The solution (with three significant digits)
\begin{eqnarray} \nonumber 
u(t)= \; &34.997\sin(0.01t-0.061)+11.323\sin(0.902t-0.777)- \nonumber \\
      &-19.489\sin(1.023t+0.054)+9.147 \sin(1.139t+0.633) \nonumber
\end{eqnarray}
is determined by the optimal vectors $\omega^\ast =(0.01, 0.902 , 1.023, 1.139)$ and $c^\ast=( 34.93, -2.151, -8.075, -7.936, 19.461, 1.048, -7.375, 5.411)$. As follows from (\ref{Eq:5}), the original problem with the solution $u(t)$ is noted as 
\begin{eqnarray} \nonumber 
\ddot{u}+10^{-4}u = & -9.213\sin(0.902t-0.777)-20.383\sin(1.023t+0.054)- \nonumber \\
                    &-11.842 \sin(1.139t+0.633), \nonumber \\
                    &u_0= -3.62, \; \dot{u}_0 = 4.556.\nonumber
\end{eqnarray}

\section{Generating a series of problems}\label{Sec:3}

The numeric experiments described below used a generator based on the approximation problem from Section \ref{Sec:2}. Apparently, the variation in any parameter of the original problem (\ref{Eq:2}) will change the optimization problem (\ref{Eq:9}), so it is sufficient to vary just a few of them.

The centers of the first two ''windows'', i.e. the pairs  $(t_1, q_1)$ and $(t_2, q_2)$, were chosen as the parameters for determining the specific problem statement. The values $q_1$ and $q_2$ were chosen independently and uniformly from the ranges  $[1,10]$ and $[-10, -1]$, respectively. The values $t_1$ and $t_2$ were dependent: first, the value $t_1$ was chosen from the range $[b+1, t_3-2]$, then, the value $t_2$ was chosen from the range $[t_1+1, t_3-1]$. All other parameters in problem (\ref{Eq:2}) were fixed: $a=1$, $b=10$, $t_3=18$, $q_0=q_3=0$, $\delta=0.3$. Parameter $\alpha$ from (\ref{Eq:8}) was chosen at $0.05$. The number of points in the additional grid $(\ref{Eq:6})$ for solving the least square problem (\ref{Eq:7}) was set at $20$. The problem of one-dimensional maximization and minimization from (\ref{Eq:9}) were solved by a scanning over a uniform grid of $100$ nodes within the interval $[a,b]$).

An important feature determining the existence of a feasible solution for the problem being considered is the number and range of frequency variation in the vector $\omega$. If the range is too small, or the number of frequencies is insufficient, the feasible domain in the problem (\ref{Eq:9}) will be empty. In the experiments carried out, the number of frequencies was chosen to be $n=3$, which corresponds to $\omega=(\omega_0,\omega_1,\omega_2,\omega_3)$, while the variable frequency change range was set from 0.01 to 2, i.e. $\omega_i\in[0.01, 2]$, $0\leq i\leq 3$.

Let's note some important properties of the problems produced by the generator under consideration.

\textbf{Remark 1}. Problem constraints (\ref{Eq:9}) are different in terms of the time required to verify them. For example, checking each of the first $n$ constraints in (\ref{Eq:9}) (let's call these constraints \textit{geometric})
\[
\omega_{i-1}(1+\alpha)-\omega_i(1-\alpha)\leq 0, \; 1\leq i \leq n,
\]
requires performing only three operations with real numbers. Testing other constraints (we will call them the \textit{main constraints}) 
\[
\left|u(t_i,\omega,c^\ast(\omega))-q_i\right|\leq\delta, \; i=1,2,3,
\]
\[
\max_{t\in[a,b]}{u(t,\omega,c^\ast(\omega))}-\min_{t\in[a,b]}{u(t,\omega,c^\ast(\omega))}\leq\delta
\]
is far more labor-intensive. First, this requires producing a system of linear equations to solve the problem (\ref{Eq:7}) ($\sim m(2n+2)^2$ computing of $\sin$ and $\cos$). Second, this system needs to be solved ($\sim \frac{2}{3}(2n+2)^3$ operations on real numbers). Third, two one-dimensional optimization problems need to be solved at the last constraint in (\ref{Eq:9}) ($\sim 100$ computing of $\sin$ and $\cos$).

\textbf{Remark 2}. The problems are characterized by multiextremal constraints, which form a non-convex feasible domain. For example, Fig.~\ref{fig:fig2} show level lines for the functions in the right-hand sides of the main constraints for one of the problems, while Fig.~\ref{fig:fig6} shows level lines of the objective function. These lines are provided within the feasible domain of the geometric constraints.
\begin{figure*}
	\centering
		\hfil\includegraphics[width=0.49\textwidth]{fig2.JPG}
		\hfil %qquad
		\includegraphics[width=0.490\textwidth]{fig3.JPG}
		\\[3ex] \hfil%qquad
		\includegraphics[width=0.490\textwidth]{fig4.JPG}
		\hfil %qquad
		\includegraphics[width=0.490\textwidth]{fig5.JPG}
		\hfil
	\caption{Level lines for the main constraints} \label{fig:fig2}	
\end{figure*}
\begin{figure}
	\centering
		\includegraphics[width=0.50\textwidth]{fig6.JPG} 
	\caption{Level lines for the objective function} \label{fig:fig6}	
\end{figure}

\textbf{Remark 3}. Increasing the frequency change range results in new solutions appearing in an area of higher frequencies; the feasible domain of the optimization problem becomes multiply connected. For example, Fig. \ref{fig:fig7} shows two trajectories, the solid line corresponds to the vector $\omega =(0.01, 0.902, 1.023, 1.139)$ and solution
\begin{eqnarray*} 
u(t)= \; &34.997\sin(0.01t-0.061)+11.323\sin(0.902t-0.777)-  \\
      &-19.489\sin(1.023t+0.054)+9.147 \sin(1.139t+0.633) 
\end{eqnarray*}
obtained at  $\omega_i\in[0.01, 2]$, $0\leq i\leq3$, dashed line corresponds to the frequency vector  $\omega=(1.749, 1.946, 2.151, 2.377)$ and solution 
\begin{eqnarray*} 
u(t)= \; &5.434 \sin( 1.749 t + 0.832 ) + 12.958 \sin( 1.946 t + 0.241)+  \\
      &+11.844 \sin( 2.151 t - 1.377 ) + 4.302 \sin( 2.377 t + 0.501)
\end{eqnarray*}
obtained at $\omega_i\in[0.01, 4]$, $0\leq i\leq 3$. Obviously, the second solution is better, as the trajectory at the last point has the largest slope.

\begin{figure}
\begin{center}
  \includegraphics[width=0.75\textwidth]{fig7.jpg} 
  \caption{Solutions with different frequencies}\label{fig:fig7} 
\end{center}
\end{figure}

These optimization problem properties allow us to conclude that this generator may be applied for testing parallel global optimization algorithms. 
%new text
%Ознакомиться с обзором подходов к распараллеливанию оптимизационных алгоритмов итересующиеся читатели могут в книге [Pardalos1992]. В рамках данного исследования мы будем применять параллельные алгоритмы, основанные на информационно-статистическом подходе [Strongin2000, 2013].
Interested readers may find the review of approaches to parallelization of optimization algorithms in \cite{Pardalos1992}. In this study we will use a parallel algorithm based on information-statistical approach \cite{Strongin2000,Sergeyev2013}.

\section{Parallel global optimization index algorithm}

Let's consider a multiextremal optimization problem in the form
\begin{equation}\label{Eq:11}
\varphi(y^\ast)=\min{\left\{\varphi(y): y\in D, g_j(y)\leq 0, 1\leq j\leq m\right\}},
\end{equation}
where the search domain $D$ is represented by a hyperparallelepiped
\begin{equation}\label{Eq:12}
D={\left\{y\in R^N: a_i\leq y_i\leq b_i, 1\leq i\leq N\right\}}.
\end{equation}
Suppose, that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_j(y), \; 1\leq j \leq m,$ of the constraints satisfy Lipschitz condition 
\[
\left|g_j(x_1)-g_j(x_2)\right|\leq L_j\left\|y_1-y_2\right\|, 1\leq j \leq m+1, y_1,y_2\in D.
\]
with respective constants $L_j, \; 1 \leq j \leq m+1,$ and may be multiextremal.
Then, it is suggested that even a single computing of a problem function value may be a time-consuming operation since it is related to the necessity of numerical modeling in the applied problems (see, for example, \cite{Grishagin2015,Modorskii}).

Using a continuous single-valued mapping$y(x)$ (Peano-type space-filling curve) of the interval  $[0,1]$ onto $D$ from (\ref{Eq:12}), a multidimensional problem (\ref{Eq:11}) can be reduced to a one-dimensional problem
\begin{equation}\label{Eq:13}
\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in [0,1], \; g_j(y(x))\leq 0, \; 1\leq j\leq m\right\}},
\end{equation}
The reduction of dimensionality matches the multidimensional problem with a Lipschitzian objective function and Lipschitzian constraints with a one-dimensional problem where the respective functions satisfy the uniform H{\"o}lder condition (see \cite{Strongin2000}), i.e. 
\[
\left|g_j(y(x_1))-g_j(y(x_2))\right|\leq K_j\left|x_1-x_2\right|^{1/N}, \;\;\; x_1,x_2\in [0,1], \;\; 1\leq j\leq m+1.
\]
Here $N$ is the dimensionality of the original multidimensional problem, and the coefficients $K_j$ are related to Lipschitz constants $L_j$ with formulae 
\[
K_j\leq 2L_j \sqrt{N+3}, \; 1\leq j\leq m+1.
\]
The issues around the numeric construction of a Peano-type curve and the corresponding theory are considered in detail in \cite{Strongin2000,Sergeyev2013}. Here we can just state that the numerically computed curve (\textit{evolvent}) is an approximation of the theoretical Peano curve with a precision at least $2^{-m}$ for each coordinate (the parameter \textit{m} is called the \textit{evolvent density}).

Let's introduce the classification of points $x$ from the search domain $[0,1]$ using the \textit{index}  $\nu=\nu(x)$. This index $\nu$ is determined by the following conditions:
\[
g_j(y(x))\leq 0, \;\;1 \leq j\leq \nu -1,\;\; g_\nu(y(x))>0,
\]
where the last inequality is negligible if $\nu=m+1$, and meets the inequalities $1\leq \nu=\nu (x) \leq m+1$. This classification produces a function
\[
f(y(x))=g_\nu (y(x)), \; \nu=\nu(x),
\]
which is determined and computed along the interval $[0,1]$. Its value in a point $x$ is either the value of the left part of the constraint violated at this point (in the case, when $\nu \leq m$), or the value of the objective function (in the case, when $\nu = m+1$). Therefore, determining the value $f(y(x))$ is reduced to a sequential computation of the values
\[
g_j(y(x)), \; 1 \leq j \leq \nu=\nu(x),
\]
i.e. the subsequent value $g_{j+1}(y(x))$ is only computed if $g_j(y(x)) \leq 0$. The computation process is completed either when the inequality $g_j(y(x))>0$ becomes true, or when the value of $\nu(x)=m+1$ is reached.

The procedure called \textit{trial} at point  $x$  automatically results in determining the index $\nu$ for this point. The pair of values  
\begin{equation}\label{Eq:16}
z=g_\nu(y(x)), \; \nu = \nu (x),
\end{equation}
produced by the trial in point $x \in [0,1]$, is called the \textit{trial result}.

A serial index algorithm for solving one-dimensional conditional optimization problems (\ref{Eq:13}) is described in detail in \cite{Barkalov2002}. This algorithm belongs to a class of characteristical algorithms (see \cite{Grishagin1997}). It can be parallelized using the approach described in \cite{Grishagin1997} for solving unconstrained global optimization problems. Let's briefly describe the rules of the resulting \textit{parallel index algorithm} (PIA).

Suppose we have $p \geq 1$  computational elements (e.g., CPU cores), which can be used to run $p$ trials simultaneously. In the first iteration of the method, $p$ trials are run in parallel at various random points $x^i\in(0,1)$, $1\leq i \leq p$. 
Suppose $n \geq 1$  iterations of the method have been completed, and as a result of which, trials were carried out in $k=k(n)$ points $x^i, 1\leq i \leq k$. Then the points $x^{k+1},...,x^{k+p}$  of the search trials in the next $(n+1)$-th iteration will be determined according to the rules below.

\begin{enumerate}
\item 
Renumber the points $x^1,...,x^k$ from previous iterations with lower indices, lowest to highest coordinate values, i.e.
\begin{equation}\label{Eq:17}
0=x_0<x_1<...<x_i<...<x_k<x_{k+1}=1,
\end{equation}
and match them with the values $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, from (\ref{Eq:16}) , calculated at these points; points $x_0=0$ и $x_{k+1}=1$ are introduced additionally; the values $z_0$ и $z_{k+1}$ are indeterminate.
\item
Classify the numbers $i,1\leq i \leq k$, of the trial points from (\ref{Eq:17}) by the number of problem constraints fulfilled at these points, by building the sets
\begin{equation}\label{Eq:18}
I_\nu = \left\{i: 1 \leq i \leq k,\ \nu = \nu(x_i)\right\},\ 1 \leq \nu \leq m+1,
\end{equation}
containing the numbers of all points $x_i,1\leq i \leq k$, with the same values of $\nu$. The end points $x_0=0$ and $x_{k+1}=1$ are interpreted as those with zero indices, and they are matched to an additional set $I_0={0,k+1}$.

Identify the maximum current value of the index
\begin{equation}\label{Eq:19}
M=\max \left\{\nu = \nu(x_i), \ 1\leq i \leq k\right\}.
\end{equation}
\item
For all values of $\nu, \ 1\leq \nu \leq m+1$, calculate the values  
\begin{equation}\label{Eq:20}
\mu_\nu = \max \left\{ \frac{\left|z_i-z_j\right|}{\left(x_i-x_j\right)^{1/N}} : i,j \in I_\nu, j<i\right\}.
\end{equation}
If the set $I_\nu$ contains less than two elements or $\mu_\nu$ from (\ref{Eq:20}) equals zero, then assume $\mu_\nu=1$.
\item
For all non-empty sets $I_\nu$, $1 \leq \nu \leq m+1$, determine the values
\begin{equation}\label{Eq:21}
  z^\ast_\nu =  
   \begin{cases}
    -\epsilon_\nu,  \nu < M, \\
    \min{\left\{g_\nu(x_i):i\in I_\nu\right\}}, \nu = M,
   \end{cases}
\end{equation}
where $M$ is the maximum current value of the index, and the vector
\begin{equation}\label{Eq:22}
\epsilon _R=\left(\epsilon_1,...,\epsilon_m\right),
\end{equation}
with positive coordinates is called the \textit{reserve vector} and is used as a parameter in the algorithm.
\item
For each interval $(x_{i-1},x_i)$,$1 \leq i \leq k+1$, calculate the \textit{characteristic} $R(i)$: 
\[
R(i)=\Delta_i+ \frac{(z_i-z_{i-1})^2}{(r_\nu\mu_\nu)^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu=\nu(x_{i-1})=\nu(x_i),
\]
\[
R(i)= 2\Delta_i-4\frac{z_i-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu(x_{i-1})<\nu(x_i)=\nu,
\]
\[
R(i)= 2\Delta_i-4\frac{z_{i-1}-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu = \nu(x_{i-1})>\nu(x_i).
\]
where $\Delta_i=(x_i-x_{i-1})^{1/N}$, and the values $r_\nu>1, 1\leq\nu\leq m+1$, are used as parameters in the algorithm.
\item
Reorder the characteristics $R(i)$, $1\leq i \leq k+1$, from highest to lowest 	
\begin{equation}\label{Eq:23}
R(t_1)\geq R(t_2)\geq ... \geq R(t_{k})\geq R(t_{k+1})
\end{equation}
and choose $p$ largest characteristics with interval numbers $t_j, 1\leq j \leq p$.
\item
Carry out $p$ new trials in parallel at the points $x^{k+j}, 1 \leq j \leq p$, calculated by the formulae
\begin{eqnarray*}
& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}, \; \nu(x_{t_j-1})\neq \nu(x_{t_j}), \\
& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}- \frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \; \nu(x_{t_j-1})=\nu(x_{t_j})=\nu. \\
\end{eqnarray*}

%Добавить результаты испытаний в информационную базу алгоритма, и перейти к правилу 1.
\end{enumerate}

The algorithm stops if the condition $\Delta_{t_j}\leq \epsilon$ becomes true for at least one number $t_j, 1\leq j \leq p$; here  $\epsilon>0$ has an order of magnitude of the desired coordinate accuracy.

Let's formulate the conditions for algorithm convergence. For this, in addition to the exact solution $y^\ast$ of the problem (\ref{Eq:11}), we will also consider the \textit{$\epsilon$-reserved solution}, determined by the conditions
\[
\varphi(y_\epsilon)=\min{\left\{\varphi(y): y\in D, \; g_j(y)\leq -\epsilon _j, \; 1\leq j\leq m\right\}},
\]
where $\epsilon_1,...,\epsilon_m$ are positive numbers  (''reserves'' for each constraint). Let's also introduce the set
\begin{equation}\label{Eq:27}
Y_\epsilon=\left\{y\in D: g_j(y)\leq 0, \; \varphi(y)\leq\varphi(y_\epsilon)\right\}
\end{equation}
of all feasible points for the problem (\ref{Eq:11}), which are no worse (in terms of the objective function's value) than  $\epsilon$-reserved solution.

Using this notation, the convergence conditions can be formulated as the theorem below.

\textbf{Theorem}. Suppose the following conditions are true:
\begin{enumerate}
\item
The problem  (\ref{Eq:11}) has $\epsilon$-reserved solution $y_\epsilon$.
\item
Each function  $g_j(y)$, $1\leq j \leq m+1$, satisfies Lipschitz condition with the respective constant $L_j$.
\item
The parameters  $\epsilon_j, \; 1 \leq j \leq m,$ from  (\ref{Eq:22}) have the values of the respective coordinates from the reserve vector $\epsilon_R$.
\item
For the values  $\mu_\nu$ from (\ref{Eq:20}) starting from a certain iteration, the following inequalities are true:
\[
r_\nu\mu_\nu>2^{3-1/N}L_\nu\sqrt{N+3}, \ 1 \leq \nu \leq m+1.
\]
\item
Each iteration uses a fixed number of computational elements  $p$, $1<p<\infty$, and each trial is completed by a single computational element within a finite time.
\end{enumerate}

Then any accumulation point $\bar{y}$ of the sequence $\left\{y^k\right\}$ generated by parallel index method while solving the problem (\ref{Eq:11}) is admissible and satisfies the conditions 
\[
\varphi(\bar{y})=\inf \left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k=1,2,...\right\}\leq \varphi (y_\epsilon).
\]
i.e., $\bar{y}$ belongs to the set $Y_\epsilon$ from (\ref{Eq:27}).

These convergence conditions proceed from the theorem of convergence of the serial index algorithm \cite{Strongin2000} and the theorem of convergence of a synchronous characteristical algorithm  \cite{Grishagin1997}.

\section{Results of numerical experiments}

%new text
The procedure applied to evaluate the efficiency of the algorithm uses an \textit{operating characteristics} method, originally  described in \cite{Grishagin1978}, which consists of the following.

Suppose a problem from the series under consideration be solved by a certain algorithm. The problem is associated with the sequence of trial points $\left\{y(x^k)\right\}$ produced by the algorithm. The sequence is truncated either when a trial point falls (for the first time) into a certain $\epsilon$-vicinity of the solution $y^*$ or when a certain number of trials $K$ does not produce such a point. In our experiments we used $K=10^6$.

The results obtained by solving all problems in a series by means of the algorithm is represented by the function $P(k)$ defined as the fraction of problems in which some trial point fall within the given  $\epsilon$-neighborhood of the solution in the first $k$ steps. This function is called the \textit{operating characteristic} of the algorithm.

Since the specification of an $\epsilon$-vicinity requires that $y^*$ be known, its value was estimated in advance (for each problem) by searching through all nodes of a uniform grid (defined on the search domain), e.g. with a step $10^{-2}$ by coordinates.

As discussed above (see Remark 1 from Section \ref{Sec:3}), the constraints in the problem being addressed have a different nature. The first three constraints are geometric, and checking whether they are feasible is not hard. Checking other constraints is far more labor-intensive. Therefore, when building operating characteristics for this class of problems, we will only consider the trials that resulted in checking labor-intensive constraints. The trials that were completed while checking geometric constraints are not included in the total number of trials.

The experiments were carried out for the parallel index algorithm (PIA) described above, with the number of used cores $p$ varying from 1 to 8. The parameters used were $r_\nu=2.5, 1\leq\nu\leq 8$. Components of the reserve vector from (\ref{Eq:22})  were selected adaptively under the rule $\epsilon_\nu=0.005\mu_\nu, 1\leq\nu\leq 7,$ where $\mu_\nu$ is from (\ref{Eq:20}) (see \cite{Strongin2000} for a justification of this component selection for the reserve vector). Computing experiments were carried out on one of the nodes of a high-performance cluster of the Nizhny Novgorod State University. The cluster node includes Intel Sandy Bridge E5-2660 2.2 GHz CPU and 64 Gb RAM. For implementation of the parallel algorithm OpenMP was used.

The algorithm's operating characteristics when using different number of computing cores $p$ obtained with a series of $100$ problems, are shown in Fig.~\ref{fig:fig8}. The location of the curves shows that any of the parallel algorithms solve 100\% of the problems in less than $7 \cdot 10^5$ trials, with more than 80\% of the problems completing within $2 \cdot 10^5$  trials. The operating characteristics also show that the algorithm is \textit{non-redundant} -- the number of trials in the parallel algorithm does not grow (compared to serial algorithm) when additional cores are employed. 
\begin{figure}[h]
\begin{center}
  \includegraphics[width=0.9\textwidth]{fig8.jpg} 
  \caption{Operating characteristics of PIA, which uses $p$ cores }\label{fig:fig8} 
\end{center}
\end{figure}

Now let's evaluate the speedup achieved by using a parallel index algorithm, depending on the number $p$ of computing cores used. Table~\ref{tab:comparison} shows the average number of iterations $n(p)$ performed by the algorithm when solving a series of $100$ problems, and speedup by the iterations $s(p)$ of a parallel algorithm. 
\begin{table}
	\caption{Speedup of the algorithm}
	\label{tab:comparison}
	\center
	\begin{tabular}{ccc}
		\hline\noalign{\smallskip}
	   $p$ & $n(p)$ & $s(p)$  \\
	  \noalign{\smallskip} \hline \noalign{\smallskip}		
	   1 &	241239 &	-- \\
		 2 &	94064 &	2.56 \\
		 4 &	45805 &	5.27 \\
		 8 &	22628 &	10.66 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}
The results show that the speedup is greater than the number of cores used (hyper-acceleration). This situation is explained by the fact that the algorithm performs an adaptive evaluation of the behavior of the objective function (calculating the lower bounds for the Lipschitz constant (\ref{Eq:20}) and the current minimum value (\ref{Eq:21})). For example, if the Lipschitz constant is better estimated in a parallel version, then the parallel algorithm using $p$ cores can be accelerated by more than $p$ times.

\section {Conclusion}

In summary, we must note that the method proposed in this work to generate multidimensional conditional global optimization problems allows:

\begin{itemize}
	\item clear visualization of the best current estimate and the final solution to the problem, regardless of the number of variables;
	\item increased dimensionality of the optimization problem being addressed by varying the original approximation problem;
	\item control of the feasible domain by adding extra non-convex constraints.
\end{itemize}
The functions included in the optimization problem are computationally intensive, which also differentiates the mechanism proposed from other known mechanisms.

The proposed generator was used to build and subsequently solve 100 problems using a parallel index algorithm. The operating characteristics of the parallel algorithm have been built, clearly demonstrating its non-redundancy.

\bigskip \textbf{Acknowledgments.} This study was supported by the Russian Science Foundation, project No 16-11-10150.

\begin{thebibliography}{99}

\bibitem{Hill}
Hill, J.D.: A search technique for multimodal surfaces. IEEE Transactions on Systems Science and Cybernetics. 5(1), 2--8 (1969)

\bibitem{Toropov}
Toropov, V.V.: Simulation approach to structural optimization. Structural Optimization. 1, 37 -- 46 (1989)

\bibitem{Shekel}
Shekel J.: Test functions for multimodal search technique. Proceedings of the 5th Princeton Conference on Information Science Systems. Princeton, Princeton University Press. 354--359 (1971)

\bibitem{Globalizer}
Strongin, R.G., Gergel, V.P., and Tropichev, A.V.: Globalizer. Investigation of minimizing sequences generated by global search algorithms for univariate functions. User's guide. Nizhny Novgorod University Press., Nizhny Novgorod (1995)

\bibitem{Strongin2000}
Strongin, R.G., Sergeyev, Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000)

\bibitem{Barkalov2002}
Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Computational Mathematics and Mathematical Physics 42(9), 1289--1300 (2002)

\bibitem{Grishagin1997}
Grishagin, V.A., Sergeyev, Ya.D., Strongin, R.G.: Parallel characteristical algorithms for solving problems of global optimization. J. Glob. Optim. 10(2), 185--206 (1997)

\bibitem{Gergel1999}
Sergeyev, Y.D., Grishagin, V.A.: Sequential and parallel algorithms for global optimization. Optimization Methods and Software. 3, 111--124 (1994)
%Gergel, V.P., Sergeyev, Ya.D.: Sequential and parallel algorithms for global minimizing functions with Lipschitzian derivatives. Computers and Mathematics with Applications, 37(4-5), 163--179 (1999)

\bibitem{Grishagin2001}
Sergeyev Ya.D, Grishagin V.A.: Parallel asynchronous global search and the nested optimization scheme. Journal of Computational Analysis and Applications. 3(2), 123--145 (2001)

\bibitem{Grishagin2016} 
Gergel, V., Grishagin, V., Gergel, A.: Adaptive nested optimization scheme for multidimensional global search. J. Glob. Optim. 66(1), 35--51 (2016)

\bibitem{Gaviano}
Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Transactions on Mathematical Software 29(4), 469--480 (2003)

\bibitem{Kvasov2006} 
Sergeyev, Ya.D., Kvasov D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim. 16(3), 910--937 (2006)

\bibitem{Sergeyev2010} Lera D., Sergeyev Ya.D. Lipschitz and Holder global optimization using space-filling curves, Applied Numerical Mathematics, 60(1-2), 115--129 (2010)

\bibitem{Paulavicius2014} 
Paulavicius, R., Sergeyev, Y., Kvasov, D., Zilinskas, J.: Globally-biased DISIMPL algorithm for expensive global optimization. J. Glob. Optim. 59(2-3), 545--567 (2014)

\bibitem{SergeyevKvasov2015} 
Sergeyev, Y.D., Kvasov, D.E.: A deterministic global optimization using smooth diagonal auxiliary functions. Communications in Nonlinear Science and Numerical Simulation. 21(1-3), 99--111 (2015)

\bibitem{Pardalos1990}
Floudas, C.A., Pardalos, P.M.: A collection of test problems for constrained global optimization algorithms. LNCS. 455 (1990)

\bibitem{Floudas1999}
Floudas, C.A., et al.: Handbook of test problems in local and global optimization.  Springer (1999)

\bibitem{Pardalos1992}
Pardalos, P.M., Phillips, A.T., Rosen, J.B.: Topics in parallel computing in mathematical programming. Science Press, New York (1992)

\bibitem{Sergeyev2013} 
Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013)

\bibitem{Grishagin2015}
Gergel, V.P., Kuzmin, M.I., Solovyov, N.A., Grishagin, V.A.: Recognition of surface defects of cold-rolling sheets based on method of localities. International Review of Automatic Control 8(1), 51--55 (2015)

\bibitem{Modorskii}
Modorskii, V.Y., Gaynutdinova, D.F., Gergel, V.P., Barkalov, K.A.: Optimization in design of scientific products for purposes of cavitation problems. In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings. 1738, art. no. 400013 (2016)

\bibitem{Grishagin1978}
Grishagin, V.A.: Operating characteristics of some global search algorithms. Problems of Stochastic Search. 7, 198--206 (1978) [in Russian]

\end{thebibliography}

\end{document}

### 2017-08-22 2017/LION_Barkalov_Lebedev/LION2017_BarkalovLebedev.tex
Added items remotely
D:\Common\Barkalov\Публикации\2017 LION\Сравнение с штрафом\LION_Barkalov_Lebedev

\documentclass{llncs}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage[misc,geometry]{ifsym}

%\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage{epstopdf}
%\usepackage{epsfig}

%\usepackage{url}
%\urldef{\mailsa}\path|i.surmin@gmail.com, avbk@mail.ru, bastrakov@vmk.unn.ru,  |
%\urldef{\mailsb}\path|evgeny.efimenko@gmail.com, arkady.gonoskov@gmail.com, meerov@vmk.unn.ru|

\begin{document}

\mainmatter

\title{Comparing Two Approaches for Solving Constrained Global Optimization Problems}
\author{Konstantin Barkalov \Letter \and Ilya Lebedev \\
\email{ \{konstantin.barkalov,ilya.lebedev\}@itmm.unn.ru}}

\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia}

\maketitle

\begin{abstract}
In the present study, a method for solving the multiextremal problems with non-convex constrains without the use of the penalty or barrier functions is considered. This \textit{index method} is featured by a separate accounting for each constraint. The check of the constraint fulfillment sequentially performed in every iteration point is terminated upon the first constraint violation occurs. The results of numerical comparing of the index method with the penalty function one are presented. The comparing has been carried out by means of the numerical solving of several hundred multidimensional multiextremal problems with non-convex constrains generated randomly.

\keywords global optimization, multiextremal functions, non-convex constraints.

\end{abstract}

\section{Introduction}

In the present paper, the methods for solving the global optimization problems with non-convex constraints
\begin{eqnarray}
&\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}, \label{problem} \\
&D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N\right\}. \label{D}
\end{eqnarray}
are considered. The objective function as well as the constraint ones are supposed to satisfy Lipschitz condition
%\[ \left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in S, \; 1\leq i \leq m+1. \]
with Lipschitz constants unknown a priori. The analytical formulae of the problem functions may be unknown, i.e. these ones may be defined by an algorithm for computing the function values in the search domain (so called ``black-box''-functions). Moreover, it is suggested that even a single computing of a problem function value may be a time-consuming operation since in the applied problems it is related to the necessity of numerical modeling (see, for example, \cite{Famularo}--\cite{Modorskii}).

The method of penalty functions is one of the most popular numerical methods for solving the problems of this kind. The idea of the method is simple and very universal; therefore, the method has found wide application to solving various practical problems. A detailed description of the method can be found, for example, in \cite{Bazaraa}. Nevertheless, the method of penalty functions has a number of disadvantages.

First, when using this method, a number of unconstrained problems with different penalty coefficients should be solved. The issue of choosing the penalty coefficients should be addressed for each problem separately.

Second, the use of large values of the penalty coefficients results in the minimization of the functions with a ravine-type minimum that essentially complicates the computations.

Third, this method is not applicable in the problems with the partially computable functions, for example, in the case when the objective function is undefined if the constraints are violated.

In the present work, the use of the penalty function method is compared to the application of the index method of accounting for the constraints developed in Nizhny Novgorod State University. The method features are (i) accounting for the information on each constraint separately and (ii) solving the problems, in which the function values may be undefined out of the feasible domain.

\section{Index method}

A novel approach to the minimization of the multiextremal functions with non-convex constraints (called the index method of accounting for the constraints) has been developed in \cite{Strongin2000}--\cite{Strongin2003}. The approach is based on a separate accounting for each constraint of the problem and is not related to the use of the penalty functions. 
At that, employing the continuous single-valued Peano curve $y(x)$ (\textit{evolvent}) mapping the unit interval $[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}) it is possible to find the minimum in (\ref{problem}) by solving a one-dimensional problem
\[
\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
\]
The considered dimensionality reduction scheme juxtaposes to a multidimensional problem with Lipschitzian functions a one-dimensional problem, where the corresponding functions satisfy uniform H{\"o}lder condition (see \cite{Strongin2000}).

According to the rules of index method, every iteration called trial at corresponding point of the search domain includes a sequential checking of fulfillment of the problem constraints at this point. The first occurrence of violation of any constraint terminates the trial and initiates the transition to the next iteration point, the values of the rest problem functions are not computed in this point. This allows: (i) accounting for the information on each constraint separately and (ii) solving the problems, in which the function values may be undefined out of the feasible domain.

The index algorithm can be applied to solving the unconstrained problems as well, i.e. in the case when $m = 0$ in the problem statement (\ref{problem}). In this case, the algorithm works analogously to its prototype -- global search algorithm (GSA) -- developed for solving the unconstrained optimization problems. Various modifications of the index algorithm and the corresponding theory of convergence are presented in \cite{Strongin2000}. The algorithm is very flexible and allows an efficient parallelization as for shared memory, as for distributed memory, as for accelerators \cite{Barkalov2010}--\cite{Lebedev2016}.

\section{Results of experiments}

Let us compare the solving of the constrained global optimization problems using penalty functions method (PM) and index method (IM). The penalty function was taken in the form 
\[
G(y) = \sigma \sum_{j=1}^m\max\left\{0,g_j(y)\right\}^2,
\]
where $\sigma$ is the penalty coefficient. Thus, a constrained problem is reduced to an unconstrained problem
\[
F(y)=\varphi(y)+G(y)
\]
which was solved using global search algorithm.

The experiments was carried out with the use of GKLS generator.  This generator for the functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.) has been proposed in \cite{Gaviano}. Eight GKLS classes of differentiable test functions of the dimensions $N = 2, 3, 4,$ and 5, have been used. For each dimension, both \textit{Hard} and \textit{Simple} classes have been considered. The difficulty of a class was increased either by decreasing the radius of the attraction region of the global minimizer, or by decreasing the distance from the global minimizer $y^\ast$ to the domain boundaries. Application of this generator for studying some unconstrained optimization algorithms has been described in \cite{Kvasov2006}--\cite{SergeyevKvasov2015}.

According to the scheme described in \cite{GergelLION}, eight series of 100 problems each based on the functions of \textit{Simple} and \textit{Hard} classes with the dimensions $N = 2, 3, 4, 5$ have been generated. There were two constraints and objective functions in each problem. The volume fraction of the feasible domain was varied from 0.2 to 0.8. The penalty parameter $\sigma$ was selected to be equal to 100.

The global minimum was considered to be found if the algorithm generates a trial point $y^k$ in the $\delta$-vicinity of the global minimizer, i.e. $\left\|y^k-y^\ast\right\|_\infty\leq\delta$. The size of the vicinity was selected as $\delta = 0.01\left\|b-a\right\|$. When using IM for \textit{Simple} class, parameter $r = 4.5$ was selected, for \textit{Hard} class $r = 5.6$ was taken. The maximum allowed number of iterations was $K_{max} = 10^6$. 
% the space-filling curve construction parameter was fixed as $m = 10$.

The average number of iterations $k_{av}$ performed by PM for solving a series of problems from both classes is shown in Table \ref{tab:PM_GKLS}. The same values for the index method are presented in Table \ref{tab:IM_GKLS}. The number of unsolved problems is specified in brackets. It reflects a situation when not all problems of the class have been solved by the method. This means that the algorithm was stopped as the maximum allowed number of iterations $K_{max}$ was achieved. In this case, the value $K_{max}=10^6$ was used for calculating the average value of the number of iterations $k_{av}$ that corresponds to the lower estimate of the average value.

\begin{table}
	\caption{The values of $k_{av}$ when solving the GKLS problems by PM}
	\label{tab:PM_GKLS}
	\center
	\begin{tabular}{cccccccccccc}
		\hline\noalign{\smallskip}
		$\Delta$ & \multicolumn{2}{c}{ $N=2$ } & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		0.2 &	1639 &	1507 & &	56149 &	60651 & & 160198(5) &	119764(2) & &	305957(1) &	488587(14) \\ 			
		0.4 & 1534 &	1967 & &	67145 &	74860 & &	158547 &	127832 & &	396480(4) &	434328(2) \\
		0.6 & 1201 &	1514 & &	92854 &	101240 & &	138550 &	143818 & &	373617(8) &	561447(9) \\
		0.8 & 1277 &	1287 & &	108121& 148479 & &	130372 &	145592 & &	488791(12) &	646538(24)\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{The values of $k_{av}$ when solving the GKLS problems by IM}
	\label{tab:IM_GKLS}
	\center
	\begin{tabular}{cccccccccccc}
		\hline\noalign{\smallskip}
		$\Delta$ & \multicolumn{2}{c}{ $N=2$ } & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		0.2 &	447 &	911 & &	14719 &	20120 & &	59680 &	66551(1) & &	391325(2) &	188797(12) \\ 			
		0.4 & 465 &	1800 & & 11951 &	17427 & &	71248 &	86899 & &	339327(1) &	151998(3) \\
		0.6 & 403 &	1988 & &	7366 &	12853 & &	58451 &	92007 & &	316648 &	179013(4) \\
		0.8 & 371	& 4292 & &	4646 &	8702 & &	33621 &	54405 & &	309844 &	124952\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Solving of GKLS problems by the index method, N=4}
	\label{tab:IM_GKLS_k}
	\center
	\begin{tabular}{cccccccc}
		\hline\noalign{\smallskip}
		$\Delta$ & \multicolumn{3}{c}{ \textit{Simple} } & & \multicolumn{3}{c}{\textit{Hard}} \\
		\noalign{\smallskip} \cline{2-4} \cline{6-8} \noalign{\smallskip}
		 & $k_1$ & $k_3$ & $k_3$& & $k_1$ & $k_2$ & $k_3$ \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		0.2 &	59680 &	20445 &	4401 & &	66551 &	24210 &	6316 \\ 			
		0.4 & 71248	& 28527	& 6784 & &	86899 &	39682 &	12615 \\
		0.6 & 58451 &	31508 &	9505 & &	92007 &	52560 &	19853 \\
		0.8 & 33621	& 21411	& 10446 & &	54405 &	36838 &	22202 \\

		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

The  average values presented demonstrate that the solving of the specified problems using the index method requires less number of iterations than with the use of the penalty function one. At the same time, separate accounting for the constraints in the index method provides less number of computations of the values of the problem functions as well. The numbers of computations of the values of the constraint functions $g_1(y), \; g_2(y)$ and those for the objective function $\varphi(y)$ ($k_1$, $k_2$, and $k_3$, respectively) are presented in Table \ref{tab:IM_GKLS_k} for four-dimensional problems.

\section {Conclusion}

Concluding, let us note that the index method for solving constrained global optimization problems considered in the present work:

\begin{itemize}
	\item is based on the global search algorithm, which is not inferior in the speed of work than other well-known algorithms;
	\item allows solving the initial problem directly, without the use of the penalty functions (thus, the issues of selection the penalty coefficient and of solving a series of unconstrained problems with different penalty coefficients are eliminated);
	\item allows solving the problems, which the values of the problem function are not defined everywhere (for example, the objective function values are undefined out of the problem feasible domain);
	\item speeds up the process of solving the constrained optimization problems (due to an essential reduction of the total number of computations of the problem function values).
\end{itemize}
The last statement has been supported by the numerical solving several hundred test problems.

\bigskip

\textbf{Acknowledgments.} This study was supported by the Russian Science Foundation, project No 16-11-10150.

\begin{thebibliography}{99}

\bibitem{Famularo}
Famularo, D., Pugliese, P., Sergeyev, Ya.D.:A global optimization technique for checking parametric robustness. Automatica. 35, 1605--1611 (1999)

\bibitem{Menniti}
Kvasov D.E., Menniti D., Pinnarelli A., Sergeyev Ya.D., Sorrentino N.: Tuning fuzzy power-system stabilizers in multi-machine systems by global optimization algorithms based on efficient domain partitions. Electric Power Systems Research. 78(7), 1217--1229 (2008)

\bibitem{Kvasov2015}
Kvasov, D.E., Sergeyev, Y.D.: Deterministic approaches for solving practical black-box global optimization problems. Advances in Engineering Software. 80, 58--66 (2015)

\bibitem{Modorskii}
Modorskii, V.Y., Gaynutdinova, D.F., Gergel, V.P., Barkalov, K.A.: Optimization in design of scientific products for purposes of cavitation problems.Solving Global Optimization Problems on GPU Cluster. In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings. 1738, art. no. 400013 (2016)

\bibitem{Bazaraa}
Bazaraa, M.S., H.D. Sherali and C. M. Shetty: Nonlinear Programming: Theory and Algorithms. Second Edition, John Wiley \& Sons, New York (1993)

\bibitem{Strongin2000}
Strongin, R.G., Sergeyev, Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000)

\bibitem{Pugliese}
Sergeyev, Ya.D., Famularo, D., Pugliese, P.: Index Branch-and-Bound Algorithm for Lipschitz univariate global optimization with multiextremal constraints. J. Glob. Optim. 21(3), 317--341 (2001)

\bibitem{Barkalov2002}
Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)

\bibitem{Strongin2003}
Strongin, R.G., Sergeyev, Ya.D.: Global optimization: fractal approach and non-redundant parallelism. J. Glob. Optim. 27(1), 25--50 (2003)

\bibitem{Barkalov2010}
Barkalov, K., Ryabov, V., Sidorov, S.: Parallel Scalable Algorithms with Mixed Local-Global Strategy for Global Optimization Problems. In: Malyshkin, V. (Ed.) MTPP 2010, LNCS. 6083, 232--240 (2010)

\bibitem{Barkalov2014}  
Barkalov, K.A., Gergel, V.P.: Multilevel scheme of dimensionality reduction for parallel global search algorithms. In: Proceedings of the 1st International Conference on Engineering and Applied Sciences Optimization -- OPT-i 2014. pp. 2111--2124 (2014)

\bibitem{Barkalov2015}
Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi Coprocessor for Solving Global Optimization Problems. In: Malyshkin, V. (Ed.) PaCT 2015, LNCS. 9251, 307--318 (2015)

\bibitem{Barkalov2016}
Barkalov, K., Gergel, V.: Parallel Global Optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016)

\bibitem{Lebedev2016}
Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster. In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings. 1738, art. no. 400006 (2016)

\bibitem{Grishagin2016} 
Gergel, V., Grishagin, V., Gergel, A.: Adaptive nested optimization scheme for multidimensional global search. J. Glob. Optim. 66 (1), 35--51 (2016)

\bibitem{Gaviano}
Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Transactions on Mathematical Software 29(4), 469--480 (2003)

\bibitem{Kvasov2006} 
Sergeyev, Ya.D., Kvasov D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim. 16(3), 910--937 (2006)

\bibitem{Paulavicius2014} 
Paulavicius, R., Sergeyev, Y., Kvasov, D., Zilinskas, J.: Globally-biased DISIMPL algorithm for expensive global optimization. J. Glob. Optim. 59(2-3), 545--567 (2014)

\bibitem{SergeyevKvasov2015} 
Sergeyev, Y.D., Kvasov, D.E.: A deterministic global optimization using smooth diagonal auxiliary functions. Commun. Nonlinear Sci. Numer. Simulat. 21(1-3), 99--111 (2015)

\bibitem{GergelLION} 
Gergel, V.: An Approach for Generating Test Problems of Constrained Global Optimization. Proceedings of the LION11 conference (2017) [submitted]

\end{thebibliography}

\end{document}

### 2017-08-22 LION_Gergel/LION2017_gergel.tex
Added items remotely
D:\Common\Barkalov\Публикации\2017 LION\Сравнение с штрафом\LION_Gergel

\documentclass{llncs}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage[misc,geometry]{ifsym}

\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage{epstopdf}
%\usepackage{epsfig}

%\usepackage{url}
%\urldef{\mailsa}\path|i.surmin@gmail.com, avbk@mail.ru, bastrakov@vmk.unn.ru,  |
%\urldef{\mailsb}\path|evgeny.efimenko@gmail.com, arkady.gonoskov@gmail.com, meerov@vmk.unn.ru|

\begin{document}

\mainmatter

\title{An Approach for Generating Test Problems of Constrained Global Optimization}
\author{Victor Gergel \\
\email{gergel@unn.ru}}

\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia}

\maketitle

\begin{abstract}
In the present paper, a novel approach to the constructing of the test global optimization problems with non-convex constraints is considered. The proposed approach is featured by a capability to construct the sets of such problems for carrying out multiple computational experiments in order to obtain a reliable evaluation of the efficiency of the optimization algorithms. When generating the test problems, the necessary number of constraints and desired fraction of the feasible domain relative to the whole global search domain can be specified. The locations of the global minimizers in the generated problems are known a priori that simplifies the evaluation of the results of the computational experiments essentially. A demonstration of the developed approach in the application to well-known index method for solving complex multiextremal optimization problems with non-convex constraints is presented.

\keywords global optimization, multiextremal functions, non-convex constraints, test optimization problems, numerical experiments

\end{abstract}

\section{Introduction}

In the present paper, the methods for generating the global optimization test problems with non-convex constraints
\begin{eqnarray}
&\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}, \label{i_problem} \\
&D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\} \label{D}
\end{eqnarray}
are considered. The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints are supposed to satisfy the Lipschitz condition
\[ \left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq i \leq m+1. \]
with the Lipschitz constants unknown a priori. The analytical formulae of the problem functions may be unknown, i.e. these ones may be defined by an algorithm for computing the function values in the search domain (so called ''black-box''-functions). It is supposed that even a single computing of a problem function value may be a time-consuming operation since it is related to the necessity of numerical modeling in the applied problems (see, for example, \cite{Famularo}--\cite{Modorskii}).

The evaluation of efficiency of the developed methods is one of the key problems in the optimization theory and applications. Unfortunately, it is difficult to obtain any theoretical estimates in many cases. As a result, the comparison of the methods is performed by carrying out the computational experiments on solving some test optimization problems in most cases. In order to obtain a reliable evaluation of the efficiency of the methods, the sets of test problems should be diverse and representative enough. The problem of choice of the test problems has been considered in a lot of works (see, for example, \cite{Floudas}--\cite{Addis}). Unfortunately, in many cases, the proposed sets contain a small number of test problems, and it is difficult to obtain the problems with desired properties. And, the most important, the constraints are absent in the proposed test problems as a rule (or the constraints are relatively simple: linear, convex, etc.).

A novel approach to the generation of any amount of the global optimization problems with non-convex constraints for performing multiple computational experiments in order to obtain a reliable evaluation of the efficiency of the developed optimization algorithms has been proposed. When generating the test problems, the necessary number of constraints and desired fraction of the feasible domain relative to the whole search domain can be specified. In addition, the locations of the global minimizers in the generated problems are known a priori that simplifies the evaluation of the results of the computational experiments essentially.

\section{Test problem classes}
A well-known approach to investigating and comparing the multiextremal optimization algorithms is based on testing these methods by solving a set of problems, chosen randomly from some specially designed class.

One generator for random samples of two-dimensional test functions has been described in \cite{Grishagin1978,Grishagin2016}. This generator produces two-dimensional functions according to the formula
\begin{eqnarray} \nonumber \label{VAG}
\varphi(y)= -&\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2+\right. \\
&\left.\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},\\ \nonumber
\end{eqnarray}
where $g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2)$, $h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2)$, $y=(y_1,y_2)\in R^2, \; 0 \leq y_1,y_2 \leq 1$, and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.

Let us consider a scheme for constructing the generator GCGen (Global Constrained optimization problem Generator) which allows to generate the test global optimization problems with $m$ constraints. Obviously, one can generate $m+1$ functions, the first $m$ of these ones can be considered as the constraints and the $(m+1)$-th function -- as the objective one. However, in this case, the conditional global minimizer of the objective function is unknown, and the preliminary estimate of this one (for example, by scanning over a uniform grid) will be time-consuming. At the same time, one could not control the size of the feasible domain. In particular, the constraints might be incompatible, and the feasible domain might be empty.

Below, the rules, which allow formulating the constrained global optimization problems so that:
\begin{itemize}
	\item one could control the size of feasible domain with respect to the whole domain of the parameters' variation;
	\item the global minimizer of the objective function would be known a priori taking into account the constraints;
	\item the global minimizer of the objective function without accounting for the constraints would be out of the feasible domain (with the purpose of simulating the behavior of the constraints and the objective function in the applied constrained optimization problems)
\end{itemize}
are proposed.

The rules defining the operation of the generator of the constrained global optimization with the properties listed above consist in the following.

\begin{enumerate}
	\item Let us generate $m+1$ functions $f_j(y), y \in D, 1 \leq j \leq m+1,$ by some generating scheme (for instance, by using the formula (\ref{VAG})). The constraints will be constructed on the base of the first $m$ functions, the $(m+1)$-th function will serve for the construction of the objective function.
	\item In order to know the global minimizer in the constrained problem a priori, let us make it to be the same to the global minimizer in the unconstrained problem. To do so, let us perform a linear transformation of coordinates so that the global minimizers of the constraint functions $y_j^\ast, \; 1 \leq j \leq m,$ would transit into the minimizer of the objective function $y_{m+1}^\ast$. This way, the functions $\overline{f_j}(y), \; 1 \leq j \leq m,$ with the same point of extremum will be constructed.
		\item In order to control the size of the feasible domain, let us construct an auxiliary function (a combined constraint)
		\[
		H(y)=\max_{1 \leq j \leq m}{\overline{f_j}(y)}
		\]
and compute its values in the nodes of a uniform grid in the domain $D$; the number of the grid nodes in the conducted experiments should be big enough (in our experiments it was $\min \left\{10^7, 10^{2N} \right\}$). Then, let us find the maximum and minimum values of the function $H(y)$ in the grid nodes, $H_{max}$ and $H_{min}$, respectively, and construct a characteristic $s(i)$ -- the number of points, in which the values of $H(y)$ fall into the range 
\[
\left[H_{min},H_{min}+i\frac{H_{max}-H_{min}}{100}\right], \; 1 \leq i \leq 100 .
\]
Then, the functions 
\[
\overline{f_j}(y) \leq q = H_{min}+i\frac{H_{max}-H_{min}}{100}, \; 1 \leq j \leq m,
\]
where $i$ is selected to be the minimal one satisfying the inequality
\[
\Delta \leq \frac{s(i)}{s(100)}
\]
will construct a problem with the feasible domain occupying the fraction $\Delta, \; 0<\Delta<1,$ of the whole search domain.
\item 
The test problem of global constrained optimization can be stated as follows
\[
\min\left\{ \varphi(y):y \in D, \; g_j(y)= \overline{f_j}(y)-q \leq 0, \; 1 \leq j \leq m \right\}
\]
where 
%$\varphi(y)=f_{m+1}(y)$. This problem will possess all declared properties (the fraction of the feasible domain will consist $\Delta$ of the whole search domain and the global minimizer of the constrained problem will be known).
%\item 
%In order to complicate the test constrained problem the form of the objective function can be expanded by the expression
\[
\varphi(y) = f_{m+1}(y)-\beta\sum_{j=1}^m\max\left\{0,\overline{f_j}(y) -q \right\}^\alpha,
\]
where $\alpha$ is a positive integer number and $\beta>0$ is selected in such a way as to provide the global minimum location of the function $\varphi(y)$ in the infeasible part of the search domain $D$. For instance, to guarantee this property the value of $\beta$ can be set as follows
\[
\beta>(h_{max}-h_{min})/(H_{max}-q)^\alpha,
\]
where $h_{max}$ and $h_{min}$ are the maximum and minimum values of the function $f_{m+1}(y)$ respectively.

\end{enumerate}

\section{Some numerical results}

As an illustration, the level lines of the objective functions and the zero-level lines of the constraints for problems constructed on the base of functions (\ref{VAG}) with $\alpha = 3, \beta = 1$ and the volume fractions of the feasible domains $\Delta = 0.4, 0.6$ are shown in Fig.~\ref{fig:VAG}. The feasible domains are highlighted by green. The change of volume and, at the same time, the increase of complexity of the feasible domains are seen clearly. Fig.~\ref{fig:VAG} (a,b) also shows the points of 628 and 764 trials, correspondingly, performed by the \textit{index method} for solving constrained global optimization problems until the required accuracy $\epsilon=10^{-2}$ was achieved. The conditional optimizer is shown as a red point and the best estimation of the optimizer is shown as a blue point.

The index method has been proposed and developed in \cite{Strongin2000,Barkalov2002}. The approach is based on a separate accounting for each constraint of the problem and is not related to the use of the penalty functions. According to the rules of index method, every iteration includes a sequential checking of fulfillment of the problem constraints at this point. The first occurrence of violation of any constraint terminates the trial and initiates the transition to the next iteration. This allows: (i) accounting for the information on each constraint separately and (ii) solving the problems, in which the function values may be undefined out of the feasible domain.  It should be noted, that the index method can be efficiently parallelized for accelerators \cite{Barkalov2015,Barkalov2016}.

\begin{figure}
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{vag_04.png} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{vag_06.png} \\ (b)}
\end{minipage}
\caption{The problems based on functions (\ref{VAG}) }
\label{fig:VAG}
\end{figure}

\section {Conclusion}

This paper considers the method for generating global optimization test problems with non-convex constraints that allows:
\begin{itemize}
	\item to control the size of feasible domain with respect to the whole domain of the parameters' variation;
	\item to known a priori the conditional global minimizer of the objective function;
	\item to generate the unconditional global minimizer of the objective function out of the feasible domain (to simulate the constraints and objective function in the applied optimization problems).
\end{itemize}

The demonstration of the developed approach in application to well-known index method for solving complex multiextremal optimization problems with non-convex constraints is considered.

The developed approach allows generating any number of test global optimization problems with non-convex constraints for performing multiple computational experiments in order to obtain a reliable evaluation of the efficiency of the developed optimization algorithms. To develop the proposed approach further, the development of new test classes for the optimization problems of various dimensionalities is planned.

\bigskip

\textbf{Acknowledgments.} This study was supported by the Russian Science Foundation, project No 16-11-10150.

\begin{thebibliography}{99}

\bibitem{Famularo}
Famularo, D., Pugliese, P., Sergeyev, Ya.D.:A global optimization technique for checking parametric robustness. Automatica. 35, 1605--1611 (1999)

\bibitem{Menniti}
Kvasov, D.E., Menniti, D., Pinnarelli, A., Sergeyev, Ya.D., Sorrentino, N.: Tuning fuzzy power-system stabilizers in multi-machine systems by global optimization algorithms based on efficient domain partitions. Electric Power Systems Research. 78(7), 1217--1229 (2008)

\bibitem{Kvasov2015}
Kvasov, D.E., Sergeyev, Y.D.: Deterministic approaches for solving practical black-box global optimization problems. Advances in Engineering Software. 80, 58--66 (2015)

\bibitem{Modorskii}
Modorskii, V.Y., Gaynutdinova, D.F., Gergel, V.P., Barkalov, K.A.: Optimization in design of scientific products for purposes of cavitation problems.Solving Global Optimization Problems on GPU Cluster. In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings. 1738, art. no. 400013 (2016)

\bibitem{Floudas} 
Floudas, C.A., et al.: Handbook of test problems in local and global optimization. Kluwer Academic Publishers, Dordrecht (1999)

\bibitem{Gaviano}
Gaviano, M., Kvasov, D.E., Lera, D., Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM TOMS 29(4), 469--480 (2003)

\bibitem{Ali} Ali, M.M., Khompatraporn, C., Zabinsky Z.B.: A numerical evaluation of several stochastic algorithms on selected continuous global optimization test problems. J. Glob. Optim. 31(4), 635--672 (2005)

\bibitem{Addis} Addis, B., Locatelli, M.: A new class of test functions for global optimization. J. Glob. Optim. 38(3), 479--501 (2007)

\bibitem{Grishagin1978}
Grishagin, V.A.: Operating Characteristics of Some Global Search Algorithms. Problems of Statistical Optimization. 7, 198--206 (1978) [in Russian]

\bibitem{Grishagin2016} 
Gergel, V., Grishagin, V., Gergel, A.: Adaptive nested optimization scheme for multidimensional global search. J. Glob. Optim. 66 (1), 35--51 (2016)

\bibitem{Strongin2000}
Strongin, R.G., Sergeyev, Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000)

\bibitem{Pugliese}
Sergeyev, Ya.D., Famularo, D., Pugliese, P.: Index Branch-and-Bound Algorithm for Lipschitz univariate global optimization with multiextremal constraints. J. Glob. Optim. 21(3), 317--341 (2001)

\bibitem{Barkalov2002}
Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)

%\bibitem{Strongin2003} Strongin, R.G., Sergeyev, Ya.D.: Global optimization: fractal approach and non-redundant parallelism. J. Glob. Optim. 27(1), 25--50 (2003)

%\bibitem{Barkalov2010} Barkalov, K., Ryabov, V., Sidorov, S.: Parallel Scalable Algorithms with Mixed Local-Global Strategy for Global Optimization Problems. In: Malyshkin, V. (Ed.) MTPP 2010, LNCS. 6083, 232--240 (2010)

%\bibitem{Barkalov2014} Barkalov, K.A., Gergel, V.P.: Multilevel scheme of dimensionality reduction for parallel global search algorithms. In: Proceedings of the 1st International Conference on Engineering and Applied Sciences Optimization -- OPT-i 2014. pp. 2111--2124 (2014)

\bibitem{Barkalov2015}
Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi Coprocessor for Solving Global Optimization Problems. In: Malyshkin, V. (Ed.) PaCT 2015, LNCS. 9251, 307--318 (2015)

\bibitem{Barkalov2016}
Barkalov, K., Gergel, V.: Parallel Global Optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016)

%\bibitem{Lebedev2016}Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster. In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings. 1738, art. no. 400006 (2016)

\end{thebibliography}

\end{document}

### 2017-09-01 2017/ParOptBook/Chapter7.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter4.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter5.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter6.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter7.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter8.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\dedic.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\editor.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\foreword.tex
D:\Common\Barkalov\Публикации\20

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                             % not available on your system
%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}% places footnotes at page bottom
%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

\title{Software Systems for Global Optimization }
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Name of First Author and Name of Second Author}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content. The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access. This allows unregistered users to read the abstract as a teaser for the complete chapter. As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf. source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript. Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}

\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content. The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access. This allows unregistered users to read the abstract as a teaser for the complete chapter. As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf. source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript. Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}

\section{Section Heading}
Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.

%\end{document}

### 2017-09-01 2017/ParOptBook/Chapter5.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter4.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter5.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter6.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter7.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter8.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\dedic.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\editor.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\foreword.tex
D:\Common\Barkalov\Публикации\20

--- 2017-09-01 2017/ParOptBook/Chapter7.tex
@@
 
-\title{Software Systems for Global Optimization }
+\title{Sequential and parallel algorithms for multidimensional multiextremal optimization based on the nested scheme of dimensionality reduction}
 % Use \titlerunning{Short Title} for an abbreviated version of
@@
 
+Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
+Further on please use the \LaTeX\ automatism for all your cross-references and citations.
+And please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
+
 %\end{document}

### 2017-09-01 2017/ParOptBook/Chapter8.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter4.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter5.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter6.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter7.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter8.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\dedic.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\editor.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\foreword.tex
D:\Common\Barkalov\Публикации\20

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%%
%%% choose options for [] as required from the list
%%% in the Reference Guide
%%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%%\usepackage{type1cm}        % activate if the above 3 fonts are
                             %% not available on your system
%%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
%%                             % when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}% places footnotes at page bottom
%
%
%\usepackage[colorlinks=true]{hyperref}
%%\hypersetup{urlcolor=blue, citecolor=red}
%

%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

\title{Supercomputer parallel computations in solving the applied optimization problems }
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Konstantin Barkalov}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Konstantin Barkalov \at Lobachevsky State University of Nizhni Novgorod,  Nizhni Novgorod, Russia \email{konstantin.barkalov@itmm.unn.ru}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\abstract*{In this chapter, the examples of solving the applied problems in the fields of optimal design (the finding of the optimum profile of a vehicle wheel), chemical research (finding the unknown kinetic parameters in the mathematical models of the complex chemical reactions), economics (the identification of the parameters in the regional economic model) are given.}

\abstract{In this chapter, the examples of solving the applied problems in the fields of optimal design (the finding of the optimum profile of a vehicle wheel), chemical research (finding the unknown kinetic parameters in the mathematical models of the complex chemical reactions), economics (the identification of the parameters in the regional economic model) are given.}

%\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content. The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access. This allows unregistered users to read the abstract as a teaser for the complete chapter. As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf. source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript. Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}

\section{Optimization of the rail transport wheel profile}

The considered problem of the search for the optimum profile of the wheels for the rail transport (railway, subway, tram, etc.) has been solved in the framework of joint research project supported by Russian Foundation for Basic Research (project number 04-01-89002-NVO\_a) and NWO (Netherlands Organization for Scientific Research, project number 047.016.014)  ``Fast Computing in Global Optimization: Sequential and Parallel Environments'', carried out at Lobachevsky State University of Nizhni Novgorod and Delft University of Technology, the Netherlands (TU Delft).

\subsection{Problem Statement}

The problem of the rail transport wheel profile optimization is described in details in \cite{8_Markine2005,8_Markine2007}. Here, a brief description of this problem is presented only. Since the wheels are of the conical shape, the center of a mounted axle is moving along a sinusoid. This process is illustrated in Fig.~\ref{8_fig_1}). The parameters of the contact between the wheel and the rail, such as the radius of rotation, the contact angle, and the angle of inclination for the mounted axle are varied at the transverse displacement of the mounted axle relative to the rail. The relationship between these variations and the transverse position of the mounted axle is determined by the wheel and rail profiles.

\begin{figure}[t]
%\sidecaption[t]
\includegraphics[width=0.9\linewidth]{figures/8_1.png}
\caption{The displacement of the mounted axle}
\label{8_fig_1}     
\end{figure}

The radius of rotation of the wheel at the contact point is an important parameter of the contact between the wheel and the rail. Actually, the radius can be different for the right and the left wheels since the mounted axle can shift relative to the rail (the radii $r_1$ and $r_2$, respectively in Fig.~\ref{8_fig_2}).

\begin{figure}[t]
%\sidecaption[t]
\includegraphics[width=0.7\linewidth]{figures/8_2.png}
\caption{The difference in the radii of rotation}
\label{8_fig_2}     
\end{figure}

When the mounted axle is in the central position, the radii of rotation are the same for the left and right wheels, i.e., $r_1=r_2=r$. The difference in the radii of rotation for the left and right wheels can be defined as a function of the transverse displacement of the mounted axle relative to its central position $\Delta r(x)=r_1(x)-r_2(x)$.

The mathematical model for this problem developed in Technical University Delft consists in the following. The wheel profile is described by B-spline, for building of which a set of points on the edge, on the edge base, and on the rolling surface of the wheel were selected (Fig.~\ref{8_fig_3})). Positions of these points can be varied for the purpose of variation of the profile. In order to reduce the computational costs for the optimization, the positions of the points on the upper surface of the edge as well as on the conical part of the profile were fixed since these parts of the wheel profile do not contact the rails.

\begin{figure}[t]
%\sidecaption[t]
\includegraphics[width=0.7\linewidth]{figures/8_3.png}
\caption{Modeling of the wheel profile}
\label{8_fig_3}     
\end{figure}

The ordinates of the movable points of the spline $z_i$ were selected as the components of the vector of the optimization problem parameters $y$, i.e.,
\[
y=(z_1,...,z_N),
\]
and the abscissas of these points were fixed.

The number of the movable points and, consequently, the number $N$ of variables in the considered problem was equal to $11$. For each parameter   the interval $\left[–1, 1\right]$ was taken as the domain of its variation. The difference of the radii of rotation $\Delta r(x)$ was minimized. The constraints were introduced according to the stability considerations. For example, one of the constraints was imposed on the maximum angle of inclination for the mounted axle.

\subsection{The results of the experiments}

Thus, the optimization problem has $N=11$ parameters and $m=6$ constraints. The objective function as well as the constraint functions is the multiextremal ones. A one-dimensional section of the objective function is presented in Fig.~\ref{8_fig_4} for illustration.

\begin{figure}[t]
%\sidecaption[t]
\includegraphics[width=0.7\linewidth]{figures/8_4.png}
\caption{A section of the objective function }
\label{8_fig_4}     
\end{figure}

The constraints have no analytical form and defined by a MATLAB computational procedures. The computation of the values of all problem functions (objective function and constraints) at one point $y$ using Pentium IV processor with the clock frequency of 3 GHz takes about 10 sec.

Thus, having formulated the problem, let us try to estimate the resources required for its solving in dependence on the method selected. Let us take, for example, the method of full scanning a uniform grid. Having selected 10 variants for the values of each parameter (remind, that the total number of parameters equals to 11), we obtain $10^{11}$ computations at the grid nodes that requires $\sim 10^{12}$ sec., i.e., $\sim 3$ years of computations using a supercomputer with 10 000 processors. Thus, the problem is so hard to solve that the parallel computations are not only necessary but should be supplied with the efficient numerical methods to obtain the results in a reasonable time unavoidably.

The problem described above has been solved by the authors using the parallel index method with the multiple shifted evolvents on a cluster of 4 computers in Delft University of Technology. The search precision in the coordinates was $2^{-10}$ , i.e., 1024 points of each parameter value in the case of the full scanning method. The search time for the optimum estimate was 27 hours. The problem functions values have been computed $4297 + 4415 + 4236 + 4266 = 17214$ times.

The computations carried out at TU Delft for a wheel of the optimum profile demonstrated its service life to increase up to 120 000 km between the profile corrections (more than five times longer as compared to the wheels of the original profile) and the maximum permitted speed to increase from 40 up to 60 m/sec.

\section{Solving the inverse problem of chemical kinetics}

In the present section the results of joint research project carried out at Lobachevsky State University of Nizhni Novgorod and at Institute of Petrochemistry and Catalysis, Russian Academy of Sciences (IPC RAS) \cite{8_Gubaidullin2011}.

\subsection{Problem statement}

The building of the mathematical models of the complex chemical reactions implies the presence of the unknown kinetic parameters (the rate constants, the activation energies, and the frequencies of collisions between the reacting molecules), which could be found by solving the problem of the minimization of the deviations between the calculated data (\textit{the direct problem}) and the experimental ones. Thus, a problem of identifying the mathematical model (\textit{the inverse problem of chemical kinetics}) arises, which in general case is a global optimization problem.

Many problems of physics chemistry imply a considerable amount of computations, nevertheless, providing rather low precision. Investigating the inverse problems of chemical kinetics requires solving many systems of differential and algebraic equations. The kinetics of complex chemical reactions is featured by the presence of parameters varying fast and slowly (because various stages of reactions go with different rates). Therefore, solving the direct kinetic problems is complicated by the hardness of the differential equations describing the mechanisms of these reactions.

The direct kinetic problem for the isothermal non-stationary model in a closed system is Cauchy problem for a system of ordinary differential equations
\begin{equation} 
 \frac{dx_i}{dt}= F_i, i=1,..,M; \; F_i=\sum_{j=1}^{N}{S_{ij}w_j}
\end{equation}
\begin{equation} 
 w_j=k_j\prod_{i=1}^{M}{\left(x_i\right)^{\left|\alpha_{ij}\right|}}-k_{-j}\prod_{i=1}^{M}{\left(x_i\right)^{\left|\beta_{ij}\right|}}
\end{equation}
with the initial conditions $t=0$, $x_i(0)=x_i^0$, where 
\begin{itemize}
	\item $x_i$ are the concentrations of the substances (\textit{the molar fractions}) participating in the reaction;
	\item $M$ is the number of substances; 
	\item $N$ is the number of steps;
	\item $S_{ij}$ are the stoichiometric matrix; 
	\item $w_j$ are the rates of the $j$-th step, $1/{hr}$; 
	\item $k_j$, $k_{-j}$ are the reduced rate constants for the forward and reverse reactions ($1/{hr}$), respectively;
	\item $\alpha_{ij}$ are the negative elements of $S_{ij}$, and $\beta_{ij}$ are the positive elements of $S_{ij}$.
\end{itemize}

Since a part of the constants $k_j$, $k_{-j}$, as a rule, are not known, the problem of identification of the mathematical model of  the inverse kinetic problem arises. This problem is a minimization one for the function of the deviation between the calculated data and the experimental ones
\begin{equation} \label{8_chem_func}
F=\sum_{i=1}^n{\sum_{j=1}^M{\left|x_{ij}^p-x_{ij}^{\epsilon}\right|}}\rightarrow \min,
\end{equation}
where
\begin{itemize}
	\item $x_{ij}^p$ are the calculated values of the observable substances concentrations  (\textit{molar fractions});	
	\item $x_{ij}^{\epsilon}$ are the values of the observable substances concentrations measured experimentally (\textit{molar fractions});
	\item $n$ is the number of the experimental points.
\end{itemize}

To find the activation energies and the collision frequencies of the molecules reacting at the elementary step, Arrhenius equation
\[
k=Ae^{-\frac{Ea}{RT}}
\]
is used, where
\begin{itemize}
	\item $k$ is the reduced rate constant for the elementary step, $1/hr$;	
	\item $E$ is the activation energy, $J/{mol}$;
	\item $R$ is gas constant, $J/{(mol \cdot K)}$;
	\item $А$ is the collision frequency for the reacting molecules;
	\item $T$ is the temperature, $K$.	
\end{itemize}

Since in the search for the rate constants of the elementary steps, these ones may fall into the area, where the differential equations system describing the reactions may appear to be a stiff one, Mishelsen method with automatic step selection is applied to solving the direct problem \cite{8_Gubaidullin2010}. As the model output data $x_{ij}^p$ depend on the values of the constants $k_j$, $k_{-j}$ nonlinearly, problem (\ref{8_chem_func}) is a multiextremal optimization problem.

\subsection{The results of the experiments}

Traditionally, the methods based on the random search ideas had been applied in IPC RAS for solving problem (\ref{8_chem_func}). The typical computation costs were several days of continuous computations. Within the framework of the joint research project by IPC RAS and Lobachevsky State University of Nizhni Novgorod it was proposed to apply the efficient parallel global optimization methods described in Chapter 5. %Ссылка на главу!!!!

For the application of the novel identification technique the mathematical model for the reaction of the catalytical carboalumination of olefins and acetylenes with the assistance of threealkylalanes in the presence of transition metal complexes was selected. This reaction has applied in the laboratory research at IPC RAS as an efficient method for the synthesis of new Me-C (methyl-carbon), Et-C (ethyl-carbon), and C-C (carbon-carbon) bonds \cite{8_Parfenova2009}. The natural experiments of conducting the reactions of the catalytical carboalumination of olefins and acetylenes for different temperatures and fixed initial concentrations have been carried out in IPC RAS. In the experiments, the concentrations of five substances were traced and their percentage ratios were determined. A number of schemes for the mathematical description of the reaction of catalytical carboalumination of olefins and acetylenes (including the parameters to be identified) have been proposed in IPC RAS. As a priority problem of the identification of the unknown parameters, it was required to determine the rate constants for the elementary steps for the proposed reaction schemes for various temperatures as well as to find the activation energies for the elementary steps for various catalysts. These schemes and the identification problem have been described in details in \cite{8_Gubaidullin2011}. Here, note that the minimization problem arising depends on 10 parameters and is box-constrained one. The computational costs for the function value at one point are $\sim 3$ sec. (Intel Xeon 3.2 GHz). The objective function is  multiextremal that is confirmed by its two-dimensional section presented in Fig.~\ref{8_fig_5}.

\begin{figure}[t]
%\sidecaption[t]
\includegraphics[width=0.8\linewidth]{figures/8_5.png}
\caption{Two-dimensional section of the objective function}
\label{8_fig_5}     
\end{figure}

As a result of the computations, the estimates of the rate constants for the elementary steps for the 10- and 12-step reactions schemes of the carboalumination in the presence of $Cp_2 ZrCl_2$ catalyst have been obtained. The objective function value at the obtained optimizer was $F = 3.040$. It appeared to be impossible to achieve the function value close to zero because the experimental data contains the uncertainties. Nevertheless, the model identified using the parallel global optimization method allowed obtaining a good conformity between the calculated data and the experimental ones. As an illustration, the comparison of the computed and experimental concentrations of substances participating in the reactions is presented in Fig.~\ref{8_fig_6}.

\begin{figure}
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{figures/8_6a.png} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{figures/8_6b.png} \\ (b)}
\end{minipage}
\caption{The comparison of the calculated data with the experimental ones}
\label{8_fig_6}
\end{figure}

\section{Identification of the dynamic balance normative models of the regional economy}

In the present section, the results obtained within the research project supported by RFBR ``Parallel global optimization methods in the identification of the dynamic balance for the normative models of the regional economy'', project 11-07-97017, carried out jointly by the research teams from Lobachevsky State University of Nizhni Novgorod and from Dorodnicyn Computing Centre of RAS are presented \cite{8_Gergel2011}.

\subsection{Problem statement}

The regional economy (the district, region, or state, etc. level) was the object of investigations. For the mathematical description of the economical processes taking place at the regional level, a dynamic balance normative economic model has been proposed in Dorodnicyn Computing Centre of RAS [7]. This model is a general one. However, it includes a significant number of parameters  being specific to the different regions. It is possible to find these parameters (or \textit{to identify the model}) by solving the problem of minimization of the deviation of the calculated time series of the macroeconomic measures from the corresponding historic statistical data. The functions entering the problem statement are the nonlinear ones. As a result, the problem of identification of the mathematical economic model is the global optimization one.

\begin{thebibliography}{99.}

\bibitem{8_Markine2005} 
Shevtsov, I.Y., Markine, V.L., Esveld, C.: Optimal design of wheel profile for railway vehicles. Wear. \textbf{258(7-8)}, 1022--1030 (2005)

\bibitem{8_Markine2007} 
Shevtsov, I.Y., Markine, V.L., Esveld, C.: An inverse shape design method for railway wheel profiles. Structural and Multidisciplinary Optimization. \textbf{33(3)}, 243--253 (2007)

\bibitem{8_Gubaidullin2011}
Gubaidullin, I.M., Ryabov, V.V., Tikhonova, M.V.: Application of the global optimization index method to solving inverse problems of chemical kinetics. Numerical methods and programming. \textbf{12(1)}, 137--145 (2011) (in Russian)

\bibitem{8_Gubaidullin2010}
Tikhonova, M.V., Gubaidullin, I.M., Spivak, S.I.: The numerical solution of the direct chemical kinetics problem by the Rosenbrock's and Mishelsen's methods for the stiff systems of differential equations. MVMS Journal. \textbf{12(12}, 26--33 (2010) (in Russian)

\bibitem{8_Parfenova2009}
Parfenova, L.V., Gabdrakhmanov, V.Z., Khalilov, L.M., Dzhemilev, U.M.: On study of chemoselectivity of reaction of trialkylalanes with alkenes, catalyzed with $Zr$ $\pi$-complexes. Journal of Organometallic Chemistry. \textbf{694(23)}, 3725--3731 (2009)

\bibitem{8_Gergel2011}
Gergel, V.P., Gorbachev, V.A., Olenev, N.N., Ryabov, V.V., Sidorov, S.V. Parallel global optimization methods for identification of the dynamic balance normative model of regional economy. Bulletin of South Ural State University. \textbf{25(9)}, 4--15 (2011) (In Russian)

\bibitem{8_Olenev1999}
Avtukhovich, E.V., Guriev, S.M., Olenev, N.N., Petrov, A.A., Pospelov, I.G., Shananin, A.A., Chukanov, S.V.: A Mathematical Model of the Transition Economy. The Computing Centre of the Russian Academy of Sciences, Moscow (1999) (In Russian)

8. N.N. Olenev. Model of innovation potential of regional economy // Proceedings of international conference “Economy of depressed areas: problems and prospects of progress of regional economy”. Barnaul, Altai State University Press. 2007. pp.178188. (In Russian)

\end{thebibliography}

%\end{document}

### 2017-09-01 2017/ParOptBook/Introduction.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter4.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter5.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter6.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter7.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter8.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\dedic.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\editor.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\foreword.tex
D:\Common\Barkalov\Публикации\20

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                             % not available on your system
%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}% places footnotes at page bottom
%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

\title{Introduction}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Name of First Author and Name of Second Author}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content. The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access. This allows unregistered users to read the abstract as a teaser for the complete chapter. As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf. source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript. Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}

\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content. The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access. This allows unregistered users to read the abstract as a teaser for the complete chapter. As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf. source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript. Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}

\section{Section Heading}
\label{sec:1}
Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.

Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations. And please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

\section{Section Heading}
\label{sec:2}
% Always give a unique label
% and use \ref{<label>} for cross-references
% and \cite{<label>} for bibliographic references
% use \sectionmark{}
% to alter or adjust the section heading in the running head
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations.

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

Use the standard \verb|equation| environment to typeset your equations, e.g.
%
\begin{equation}
a \times b = c\;,
\end{equation}
%
however, for multiline equations we recommend to use the \verb|eqnarray|
environment\footnote{In physics texts please activate the class option
\texttt{vecphys} to depict your vectors in \textbf{\itshape
boldface-italic} type - as is customary for a wide range of physical
subjects}.
\begin{eqnarray}
a \times b = c \nonumber\\
\vec{a} \cdot \vec{b}=\vec{c}
\label{eq:01}
\end{eqnarray}

\subsection{Subsection Heading}
\label{subsec:2}
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references\index{cross-references} and citations\index{citations} as has already been described in Sect.~\ref{sec:2}.

\begin{quotation}
Please do not use quotation marks when quoting texts! Simply use the \verb|quotation| environment -- it will automatically render Springer's preferred layout.
\end{quotation}

\subsubsection{Subsubsection Heading}
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{subsec:2}, see also Fig.~\ref{fig:1}\footnote{If you copy text passages, figures, or tables from other works, you must obtain \textit{permission} from the copyright holder (usually the original publisher). Please enclose the signed permission with the manuscript. The sources\index{permission to print} must be acknowledged either in the captions, as footnotes or in a separate section of the book.}

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

% For figures use
%
\begin{figure}[b]
\sidecaption
% Use the relevant command for your figure-insertion program
% to insert the figure file.
% For example, with the graphicx style use
\includegraphics[scale=.65]{figure}
%
% If no graphics program available, insert a blank space i.e. use
%\picplace{5cm}{2cm} % Give the correct figure height and width in cm
%
\caption{If the width of the figure is less than 7.8 cm use the \texttt{sidecapion} command to flush the caption on the left side of the page. If the figure is positioned at the top of the page, align the sidecaption with the top of the figure -- to achieve this you simply need to use the optional argument \texttt{[t]} with the \texttt{sidecaption} command}
\label{fig:1}       % Give a unique label
\end{figure}

\paragraph{Paragraph Heading} %
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

For typesetting numbered lists we recommend to use the \verb|enumerate| environment -- it will automatically render Springer's preferred layout.

\begin{enumerate}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\begin{enumerate}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\end{enumerate}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\end{enumerate}

\subparagraph{Subparagraph Heading} In order to avoid simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}, see also Fig.~\ref{fig:2}.

For unnumbered list we recommend to use the \verb|itemize| environment -- it will automatically render Springer's preferred layout.

\begin{itemize}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development, cf. Table~\ref{tab:1}.}
\begin{itemize}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\end{itemize}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\end{itemize}

\begin{figure}[t]
\sidecaption[t]
% Use the relevant command for your figure-insertion program
% to insert the figure file.
% For example, with the option graphics use
\includegraphics[scale=.65]{figure}
%
% If no graphics program available, insert a blank space i.e. use
%\picplace{5cm}{2cm} % Give the correct figure height and width in cm
%
%\caption{Please write your figure caption here}
\caption{If the width of the figure is less than 7.8 cm use the \texttt{sidecapion} command to flush the caption on the left side of the page. If the figure is positioned at the top of the page, align the sidecaption with the top of the figure -- to achieve this you simply need to use the optional argument \texttt{[t]} with the \texttt{sidecaption} command}
\label{fig:2}       % Give a unique label
\end{figure}

\runinhead{Run-in Heading Boldface Version} Use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.

\subruninhead{Run-in Heading Italic Version} Use the \LaTeX\ automatism for all your cross-refer\-ences and citations as has already been described in Sect.~\ref{sec:2}\index{paragraph}.
% Use the \index{} command to code your index words
%
% For tables use
%
\begin{table}
\caption{Please write your table caption here}
\label{tab:1}       % Give a unique label
%
% Follow this input for your own table layout
%
\begin{tabular}{p{2cm}p{2.4cm}p{2cm}p{4.9cm}}
\hline\noalign{\smallskip}
Classes & Subclass & Length & Action Mechanism  \\
\noalign{\smallskip}\svhline\noalign{\smallskip}
Translation & mRNA$^a$  & 22 (19--25) & Translation repression, mRNA cleavage\\
Translation & mRNA cleavage & 21 & mRNA cleavage\\
Translation & mRNA  & 21--22 & mRNA cleavage\\
Translation & mRNA  & 24--26 & Histone and DNA Modification\\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)
\end{table}
%
\section{Section Heading}
\label{sec:3}
% Always give a unique label
% and use \ref{<label>} for cross-references
% and \cite{<label>} for bibliographic references
% use \sectionmark{}
% to alter or adjust the section heading in the running head
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

If you want to list definitions or the like we recommend to use the Springer-enhanced \verb|description| environment -- it will automatically render Springer's preferred layout.

\begin{description}[Type 1]
\item[Type 1]{That addresses central themes pertainng to migration, health, and disease. In Sect.~\ref{sec:1}, Wilson discusses the role of human migration in infectious disease distributions and patterns.}
\item[Type 2]{That addresses central themes pertainng to migration, health, and disease. In Sect.~\ref{subsec:2}, Wilson discusses the role of human migration in infectious disease distributions and patterns.}
\end{description}

\subsection{Subsection Heading} %
In order to avoid simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Use the \LaTeX\ automatism for all your cross-references and citations citations as has already been described in Sect.~\ref{sec:2}.

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

\begin{svgraybox}
If you want to emphasize complete paragraphs of texts we recommend to use the newly defined Springer class option \verb|graybox| and the newly defined environment \verb|svgraybox|. This will produce a 15 percent screened box 'behind' your text.

If you want to emphasize complete paragraphs of texts we recommend to use the newly defined Springer class option and environment \verb|svgraybox|. This will produce a 15 percent screened box 'behind' your text.
\end{svgraybox}

\subsubsection{Subsubsection Heading}
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

\begin{theorem}
Theorem text goes here.
\end{theorem}
%
% or
%
\begin{definition}
Definition text goes here.
\end{definition}

\begin{proof}
%\smartqed
Proof text goes here.
\qed
\end{proof}

\paragraph{Paragraph Heading} %
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.

Note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
%
% For built-in environments use
%
\begin{theorem}
Theorem text goes here.
\end{theorem}
%
\begin{definition}
Definition text goes here.
\end{definition}
%
\begin{proof}
\smartqed
Proof text goes here.
\qed
\end{proof}
%
\begin{acknowledgement}
If you want to include acknowledgments of assistance and the like at the end of an individual chapter please use the \verb|acknowledgement| environment -- it will automatically render Springer's preferred layout.
\end{acknowledgement}
%
\section*{Appendix}
\addcontentsline{toc}{section}{Appendix}
%
%
When placed at the end of a chapter or contribution (as opposed to at the end of the book), the numbering of tables, figures, and equations in the appendix section continues on from that in the main text. Hence please \textit{do not} use the \verb|appendix| command when writing an appendix at the end of your chapter or contribution. If there is only one the appendix is designated ``Appendix'', or ``Appendix 1'', or ``Appendix 2'', etc. if there is more than one.

\begin{equation}
a \times b = c
\end{equation}

\input{referenc}
%\end{document}

### 2017-09-01 2017/ParOptBook/Chapter1.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter4.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter5.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter6.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter7.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter8.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\dedic.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\editor.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\foreword.tex
D:\Common\Barkalov\Публикации\20

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                            %% not available on your system
%%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
                             %% when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}
%\usepackage{amsmath}
%
%% places footnotes at page bottom
%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
                       %% please use the style svind.ist with
                       %% your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

\title{Global optimization: from one-dimensional to multidimensional problems via the dimensionality reduction}
\titlerunning{Global optimization via dimensionality reduction}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Vladimir A.Grishagin and Yaroslav D.Sergeyev}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Vladimir A.Grishagin \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia   \email{vagris@unn.ru}
\and Yaroslav D.Sergeyev \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia 
\at DIMES, University of Calabria, Via P. Bucci, Cubo 42-C, 87036 Rende (CS), Italy \email{yaro@si.deis.unical.it}}
%
\maketitle

\abstract{This chapter is devoted to the basic statements and general definitions concerning the global optimization problems and algorithms. At the beginning, relevant notions are introduced and different statements of global optimization problems are described. As a main optimization statement, a non-linear programming problem is considered.  The concepts of numerical methods for solving the optimization problems are discussed and a general abstract model for successive methods is introduced. On the base of abstract description its algorithmic implementation (general computational scheme) is given. The model of sequential optimization expands for the case of parallel computations and general computational procedures for synchronous and asynchronous parallel methods are presented. Convergence of optimization methods and possibility of obtaining estimations of the problem solution after a finite number of objective function evaluations are discussed. Significant complexity of multidimensional multie
xtremal optimization problems is considered and approaches of overcoming this complexity via reduction schemes are briefly described. }

\section{The global optimization problem}
\label{sec:1}
The aim of this section is to introduce the necessary terms and to formulate the general problem of the global optimization, the investigation of which the present monograph is dedicated to. 
Let $\varphi(y)$ be a real-valued function, defined over a domain $D$ of the $N$-dimensional Euclidean space $R^N$ with a finite value at every point of $Q$.

Let us designate 
\begin{equation}
\label{eq:1}
\inf_{y\in Q}\varphi(y) \equiv\inf\{\varphi(y):y \in Q\}
\end{equation}
as the infimum of $\varphi(y)$ in the domain $Q$.
\begin{definition}
If there exists a point  $y^*\in Q$  such that 
\begin{equation}
\label{eq:2}
\varphi(y^*)=\inf\{\varphi(y):y \in Q\}
\end{equation}
then $\varphi(y)$ reaches its infimum in the domain $Q$, and $y^*$ is \textit {the global (absolute) minimum point} or \textit {the global minimizer}. The value $\varphi^* = \varphi(y^*)$ is the lowest or global optimum (minimum) value of $\varphi$ in $Q$ and is denoted as 
\begin{equation}
\label{eq:3}
\min_{y\in Q}\varphi(y) \equiv\min\{\varphi(y):y \in Q\}.
\end{equation}
\end{definition}
Let us denote the set of all points $y\in Q$  satisfying (\ref{eq:2}) as
\begin{equation}
\label{eq:4}
Q^*=Arg\min_{y\in Q}\varphi(y) \equiv Arg\min\{\varphi(y):y \in Q\}.
\end{equation}

Let us give one more important definition featuring the extremal properties of the functions.
\begin{definition} 
A point $y'$  is called a \textit {local minimum point} in the domain $Q$ if there exists such number $\epsilon>0$ that for all $y \in Q$ satisfying the inequality $\left\|y-y'\right\|$ the condition $\varphi(y')<\varphi(y)$ is true.
\end{definition}
The functions with several local minima in the domain $Q$ are called the \textit {multiextremal} ones in this domain, and those with a single local minimum are called the \textit {unimodal (uniextremal)} functions. Obviously, any global minimum is a local one, and for a unimodal function its unique local minimum is the global one.
In the optimization models related to the analysis of the function $\varphi(y)$, different formalizations of the desired solution are possible and, as a result, different optimization problem statements arise. In this connection, let us give a classification of such problem statements in the following formulations.
\begin{definition} 
Let us treat the optimization problem as a problem of the following kind: to find the predefined extremal characteristics of the function $\varphi(y)$ over the domain $Q$.
\end{definition}
Various optimization problem statements are possible subject to the desired extremal characteristics.

\textbf{Problem Statement A.} Find the infimum of the function $\varphi(y)$
\begin{equation}
\label{eq:5}
\varphi^*=\inf\{\varphi(y):y \in Q\}.
\end{equation}

\textbf{Problem Statement B.} Find the infimum from (\ref{eq:5}) and if the set of the global minimum points from (\ref{eq:2}) is not empty, find at least one point $y^*=Q^*$.

\textbf{Problem Statement C.} Find the infimum from (\ref{eq:5})  and find all global minimum points (or make sure that the set $Q^*$ is empty).

\textbf{Problem Statement D.} Find all coordinates and values of all local minima of the function $\varphi(y)$ in the domain $Q$.

The problem statements A—D are related to the finding of the global minimum, but sometimes the following problem statement is used even for the multiextremal functions:

\textbf{Problem Statement E.} Find a local minimum of the function $\varphi(y)$ in the domain $Q$.

The optimization problem statements A–E are the most commonly used (see, for example, classical monographs \cite{1_FloudasPardalos, 1_Himmelblau, 1_HorstPardalos, 1_HorstTuy, 1_McCormick, 1_Mockus, 1_Pinter,  1_SergKvasMonogr, 1_StrSergMon2000, 1_ZhigZhil}), although other variants are possible as well, for example, those given in \cite{1_Himmelblau}. The desired extremal characteristics defined by the problem statement $Z\in\{A, B, C, D, E\}$ are referred to as $Z$-solution of the optimization problem.

Let us denote the general optimization problem symbolically in the following form:
\begin{equation}
\label{eq:6}
\varphi(y)\rightarrow\inf, y \in Q.
\end{equation}

This problem is rather often referred to as the problem of the \textit{mathematical programming}.

If it is necessary to concretize the problem statement to be considered, we will specify the relevant desired $Z$-solution of the problem.

As  the supremum of the function $\varphi(y)$ in the domain $Q$ is connected with the infimum of the function $-\varphi(y)$ according to  the expression
\begin{displaymath}
\sup\{\varphi(y):y \in Q\} = -\inf\{-\varphi(y):y \in Q\}
\end{displaymath}
the problem of finding the extremal characteristics related to the maximum of the function $\varphi(y)$ (the maximization problem), can be reduced to the problem of the minimization of the function $-\varphi(y)$. Therefore, hereinafter the optimization problem will be considered in the form (\ref{eq:6}). The function $-\varphi(y)$ in (\ref{eq:6}) will be called \textit{the objective function, the minimized function, or the optimized function},  the domain $Q$ will be called \textit{the feasible (admissible) domain}, and the elements of the set $Q$ will be called \textit{the feasible points}.

Let us formulate the problem (\ref{eq:6}), for which the set of the global minimum points $Q^*$ is known to be nonempty a priori (the sufficient conditions for the non-emptiness of $Q^*$ are given, for example, by Weierstrass theorem), as
\begin{equation}
\label{eq:7}
\varphi(y)\rightarrow\min, y \in Q.
\end{equation}

Note that the mathematical programming problem is often formulated just in this form, although, of course, the problem (\ref{eq:6}) is the more general one.
In the present monograph, the existence of a global minimizer is always presumed. Let us reformulate the problem (\ref{eq:8}) in the form of a non-linear programming problem
\begin{equation}
\label{eq:8}
\varphi(y)\rightarrow\min, y \in Q\subseteq R^N,
\end{equation}
\begin{equation}
\label{eq:9}
Q=\{y \in D:g_j(y)\leq 0,1\leq j\leq m \},
\end{equation}
\begin{equation}
\label{eq:10}
D=\{y \in R^N:y_i\in [a_i,b_i],1\leq i\leq N \},
\end{equation}
i.e., the problem of finding the extremal values (in terms of the Problem statements A-D) of the objective (minimized) function $\varphi(y)$  over the domain $Q$ is defined by the \textit{coordinate} (\ref{eq:10}) and \textit{functional} (\ref{eq:9}) \textit{constraints} on the selection of feasible points (or vectors) $y=(y_1,\ldots ,y_N)$ . In this model, the constants $a_i,b_i, 1\leq i\leq N$,  determine the limits for the variation of the problem parameters (the components of the vector $y$), and are finite or (when the corresponding upper or/and lower limits are absent) are assumed to be equal to $a_i=-\infty $  and/or $b_i=+\infty $ .

Often, the constraints in the form of equations are also included in the non-linear programming problem statement. However, any equation $h(y)=0$ , first, could be represented formally as the system of two inequalities $h(y)\leq 0$  and $h(y)\geq 0$ . Second, in numerical solving an optimization problem on a computer the exact calculation of the equality $h(y)=0$ is often impossible, therefore, this equality is assumed to be computed  with some error $\delta>0$ , i.e., the inequality $\left|h(y)\right|\leq \delta$ is considered instead of the equation $h(y)=0$. Thus, taking into account the above reasons, one can claim that (\ref{eq:8})-(\ref{eq:10}) is the statement of  \textit {the general  problem of the nonlinear programming}.

If $m=0$ , i.e., there are no functional constraints, let us assume $Q=D$. The problem (\ref{eq:8})-(\ref{eq:10}) in this case will be called \textit {the unconstrained optimization problem}.

Let us consider an optimization problem  for the function
\begin{displaymath}
\label{eq:12}
\varphi(y_1,y_2)=-1.5y_1^2exp(1-y_1^2-20.25(y_1-y_2)^2)-
\end{displaymath}
\begin{displaymath}
\label{eq:12}
-(0.5(y_1-1)(y_2-1))^4exp(2-(0.5(y_1-1))^4-(y_2-1)^4)
\end{displaymath}
within the domain $0 \leq y_1\leq 4,\ \  -1\leq y_2\leq 3$, with the following constraints 
\begin{displaymath}
g_1(y_1,y_2)=0.01((y_1-2.2)^2+(y_2-1.2)^2-2.25)\leq 0,
\end{displaymath}
\begin{figure}[t]
\includegraphics[width=0.7\linewidth]{figures/figure1_1.png}
\caption{Example of a global optimization problem with disconnected admissible domain}
\label{1_fig_1}    
\end{figure}
\begin{displaymath}
g_1(y_1,y_2)=100(1-(y_1-2)^2/1.44-(0.5y_2)^2)\leq 0,
\end{displaymath}
\begin{displaymath}
g_1(y_1,y_2)=10(y_2-1.5-1.5sin(6.283(y_1-1.75)))\leq 0,
\end{displaymath}
as a particular example.
The set of the points satisfying the first constraint represents a circle with the boundary defined by the equation $g_1(y_1,y_2)=0$. The points satisfying the second constraint are located outside an ellipse defined by the equation $g_2(y_1,y_2)=0$. The points satisfying the third constraint are below the sinusoid $g_3(y_1,y_2)=0$. Thus, the feasible domain is disconnect and consists of three non-convex subdomains (the feasible domain is dark in Fig.\ref{1_fig_1}). The global minimum $\varphi(y_1^*,y_2^*)=-1.489$  is achieved at the point $(y_1^*,y_2^*)$  = (0,942, 0,944).

\section{Numerical optimization methods}
\label{sec:1.2}

\subsection {Concept of algorithm. Sequential scheme}
\label {subsec:1.2.1}

Having stated the minimization problem, we now must answer the principal question: how to solve it?

The classical approach of the mathematical analysis suggests the following procedure for the analytical solution of the problem. Let us consider the one-dimensional case and assume that $\varphi(y)$ is a single argument function, which is piecewise smooth over the interval $[a,b]$. Then, the minimum of $\varphi(y)$ in $[a,b]$ can be achieved only at the points, where the derivative $\varphi'(y)=0$ , or where the derivative is discontinuous, or at the boundary points. Therefore, it is necessary to find all these points and to select from them the point with the lowest value. In other words, in order to solve the problem by this method, the following requirements should be satisfied:
\begin{enumerate}
\item{it is necessary to know the analytical formula for the function;}
\item{the function should be piecewise smooth;}
\item{the calculation of the derivative should be possible;}
\item{it is necessary to solve the equation$\varphi'(y)=0$;}
\item{the information on the derivative discontinuity points or a method for finding these ones are required.}
\end{enumerate}

Unfortunately, these requirements are fulfilled rarely. In the typical cases, the function $\varphi(y)$ is defined algorithmically, i.e., in the form of some calculation scheme, when one can calculate the value of $\varphi(y)$ at a given point y only (black-box function). In this case, the analytical investigation methods are unsuitable. Note that even if the function is defined analytically and if the derivative calculation is possible the initial problem (\ref{eq:6}) is reduced to the problem of finding all the derivative roots, the difficulty of which is comparable with the minimization problem.

The limited applicability of the analytical methods has caused the development and wide application of \textit {the numerical methods (algorithms)} for solving the optimization problems. Various definitions of the numerical optimization method have been given by many authors. The consideration of the method as an iteration procedure, which calculates certain characteristics of the minimized function at some points of the search domain is a common feature of all formulations. The calculated characteristics can include the value of the objective function, its gradient, the second derivatives matrix, etc. Let us denote the calculation of the function characteristics at a point as \textit {a search trial} and the set of the values of the function characteristics as \textit {a trial result}.

Hereinafter we will consider the function value at the trial point as the trial result. If an algorithm is executed on a single processor computer, the computing scheme of the algorithm consists of a series of operations executed sequentially, and all the trials are performed sequentially, i.e., we deal with a pure sequential numerical method.

Let us give a formal definition of sequential optimization algorithm based on the methodology of the theory of operations research, or, more general, on the computational model for solving the problem (\ref{eq:6}). The generalization of this model onto the parallel computing case will be given in Subsect. \ref{subsec:1.2.2}

The development of a computational model presupposes an availability of some a priori information on the problem to be solved (before the beginning of the computations). This information could be obtained from the physical meaning of the problem or description of the properties of the model-based object. This information can include such properties of $\varphi(y)$ as continuity, smoothness, monotonicity, convexity, etc. The available information provides the researcher with the basis to attribute the problem (in our case, the function $\varphi(y)$ to a specific set (or class) $\Phi$. After the class $\Phi$ has been identified, the information known to the researcher a priori consists in the knowledge that the problem belongs to the class $\Phi$.

The selection of an \textit {optimization algorithm} (method for solving  the optimization problem) is the next important step in the computational model development. In the most general form, a numerical method $s$ for solving the problem from the class $\Phi$ is a set  (tuple) (\cite{1_StrMonRus})
\begin{equation}
\label{eq:11}
s=\left\langle \{G_k\},\{E_k\},\{H_k\}\right\rangle
\end{equation}
where

-  $\{G_k\}$	  is a set of functionals defining the rules for the trial points selection \textit {(algorithm's decision rules)}, $k=1,2,\ldots$ ;

-  $\{E_k\}$	  is a set of functionals defining the rules of constructing the approximate solution (the estimate of extremum), $k=1,2,\ldots$ ;

-  $\{H_k\}$	  is a set of functionals defining the rules of stopping the computational process,  $k=1,2,\ldots$.

The order of the trial execution or the computational scheme of the algorithm   can be realized in accordance with the following steps.

\begin{description}[\textbf{Step 1}]
\item[\textbf{Step 1}]{The point 
\begin{equation}
\label{eq:12}
y^1=G_1(\Phi\in Q)
\end{equation}
is selected as the first trial coordinate and the trial counter $k=1$ .}
\item[\textbf{Step 2}]{Assume the point of the $k$-th trial $y^k\in Q$  $(k\geq 1)$ has been selected. The value of the function $z^k=\varphi(y^k)$  is computed. Afterwards there is the following \textit {search (a posteriori) information} about the function $\varphi(y)$:
\begin{equation}
\label{eq:13}
\omega_k=\{(y^1,z^1),(y^2,z^2),\ldots ,(y^k,z^k)\}.
\end{equation}
This information allows reducing the class, which the function  $\varphi(y)$ belongs to, up to the set
\begin{equation}
\label{eq:14}
\Phi(\omega_k)=\{\psi\in\Phi:\psi(y^i)=z^i,1\leq i \leq k\}.
\end{equation}
}
\item[\textbf{Step 3}]{The current estimate 
\begin{equation}
\label{eq:15}
e^k=E_k(\Phi,\omega_k)
\end{equation}
of the extremum sought (approximate solution) is determined.}
\item[\textbf{Step 4}]{The next trial point 
\begin{equation}
\label{eq:16}
y^{k+1}=G_{k+1}(\Phi,\omega_k)
\end{equation}
is computed.}
\item[\textbf{Step 5}]{The number 
\begin{equation}
\label{eq:17}
h^k=H_k(\Phi,\omega_k,y^{k+1})\in \{0,1\}
\end{equation}
taking one of two possible values: 0 or 1 is determined. If $h^k=1$ the trial counter $k$ is increased by 1, and the return to the Step 2 of the scheme is carried out. If $h^k=0$ the computations are terminated, and the estimate $e^k$  is taken as the problem solution.}
\end{description}

The general computation model has been described.

\example {Let us consider the simplest method of solving a one-dimensional problem (\ref{eq:6}) within an interval $[a,b]$, namely, the method of scanning  the function values in the nodes of a regular grid. The method consists in the partitioning of the search interval $[a,b]$ into $q$ equal parts. The function values are calculated at the points (nodes) of the partitioning, including the ends of the search interval. The lowest calculated value (and its coordinate, if it is required by the problem statement) is considered to be the solution of the problem. }

The computability of the function at any point of the search domain is sufficient for the applicability of the method. So, the class of the functions defined over the interval $[a,b]$ and being computable at every point of the interval could be considered as the a priori known class $\Phi$.

For this method
\begin{displaymath}
G_1(\Phi)=a, G_{k=1}(\Phi,\omega_k)=a+k\frac{b-a}{q}, k\geq 1,
\end{displaymath}
\begin{displaymath}
  H_k(\Phi,\omega_k) =
  \begin{cases}
    0, & k > q, \\
    1, & k \leq q,
  \end{cases}
\end{displaymath}
\begin{equation}
\label{eq:18}
e^k=\varphi_k^*,
\end{equation}
where
\begin{equation}
\label{eq:19}
\varphi_k^*=\min_{1\leq i\leq k} \varphi(y^i),
\end{equation}
or
\begin{displaymath}
e^k=(\varphi_k^*,y_k^*),
\end{displaymath}
where 
\begin{displaymath}
\label{eq:19}
y_k^*=arg\min_{1\leq i\leq k} \varphi(y^i).
\end{displaymath}

\subsection {Parallel optimization algorithms}
\label {subsec:1.2.2}
The computational scheme (\ref{eq:12})-(\ref{eq:17}) implies  the sequential execution of all the algorithm steps, that can be provided by the single processor computers. But if a multiprocessor computer is available, a parallelization of the computational scheme and, therefore, an acceleration of the optimization process can be realized. The parallelization of optimization methods could be done in several ways. It is possible to parallelize, in the first place, the calculation of the objective function, secondly, the implementation of the rules $G_k,E_k,H_k$ , and, finally, the scheme (\ref{eq:12})-(\ref{eq:17}) providing simultaneous execution of several trials. However, the parallelization of the objective function describing the optimized object is specific for different objective function models, and as a matter of fact, cannot be related to the general optimization scheme. The parallelization of the functionals  $G_k,E_k,H_k$ depends essentially on the class of algorithms. Besides, often these functionals are rather simple, and their parallelization is just of no use. Therefore, Therefore, the subject of the consideration in the present subsection  is the design of some \textit {fundamental principles} for a parallel execution of trials.  Let us use the term \textit {iteration} for the simultaneous (parallel) execution of several trials (each trial is executed by a separate processor). Let us denote the number of trials during the $n$-th iteration as $p(n)$ , and the total number of trials executed during all $n$ iterations as $k(n)$ . In other words, we assume that we have  $p$  processors at our disposal while executing the $n$-th iteration. Actually, the number of processors employed can be different for different iterations.

Let us consider the synchronous variant of the parallelization, where the transition to the next iteration is performed after full termination of the current one, i.e., after completing the last trial of the current iteration. In this case, it is obvious that $k(n)=p(1)+\ldots p(n),n\geq 1$.  Let us denote the vector of the trial coordinates at the $n$-th iteration as $u(n)$ , i.e.,
\begin{displaymath}
u(n)=(y^{k(n-1)+1},\ldots ,y^{k(n)}) .
\end{displaymath}
and the first $j$ coordinates of the vector $u(n)$  as $u_j(n)$ , i.e., 
\begin{displaymath}
u(n)=(y^{k(n-1)+1},\ldots ,y^{k(n-1)+j}),\ 1\leq j\leq p(n) .
\end{displaymath}

Then, the general computational scheme of a synchronous parallel optimization algorithm can be written as follows.

\paragraph{\textbf{Synchronous parallel algorithm scheme}}
\begin{description}[\textbf{Step 1}]
\item[\textbf{Step 1}]{Set the iteration number $n=1$  and the initial number of trials $k(0)=0$ . Select $p(1)$ points of the first iteration  
\begin{equation}
\label{eq:20}
y^1=G_1(\Phi\in Q),y^i=G_i(\Phi,u_{i-1}(1))\in Q,\ 2\leq i\leq p(1)
\end{equation}
and assign the number of trials $k(1)=p(1)$.}
\item[\textbf{Step 2}]{Assume that the points of the $n$th iteration $u(n),\ n\geq 1,$ , have been obtained. The computations of the objective function values $z^j=\varphi(y^j),\ k(n-1)+1\leq j\leq k(n)$, are executed (each value is computed by a separate processor). After computing all the values there is the following search information on the function $\varphi(y)$ (\textit{a posteriori information}):
\begin{equation}
\label{eq:21}
\omega_k=\omega_{k(n)}=\{(y^1,z^1),(y^2,z^2),\ldots ,(y^{k(n)},z^{k(n)})\}.
\end{equation}
As before, the new information (\ref{eq:21}) allows one to reduce the class, which the function $\varphi(y)$ belongs to, up to the set   
\begin{equation}
\label{eq:22}
\Phi(\omega_{k(n)})=\{\psi\in\Phi:\psi(y^i)=z^i,1\leq i \leq k(n)\}.
\end{equation}
}
\item[\textbf{Step 3}]{After completing all the trials within the current iteration, the estimate of the extremum (an approximate solution) is formed:  
\begin{equation}
\label{eq:23}
e^{k(n)}=E_{k(n)}(\Phi,\omega_{k(n)}).
\end{equation}
}
\item[\textbf{Step 4}]{$p(n+1)$   coordinates of the trials of the $n+1$-th iteration are computed:
\begin{equation}
\label{eq:24}
\begin{gathered}
y^{k(n)+1}=G_{k(n)+1}(\Phi,\omega_{k(n)}) \\
y^i=G_i(\Phi,\omega_{k(n)},u_{i-1}(k(n+1))
\end{gathered}
\end{equation}
and the number of the search trials $k(n+1)=k(n)+p(n+1)$  is set.}
\item[\textbf{Step 5}]{The number 
\begin{equation}
\label{eq:25}
h^{k(n)}=H_{k(n)}(\Phi,\omega_{k(n)},u_{n+1})\in \{0,1\}
\end{equation}
taking one of two possible values: 0 or 1 is determined. If $h^{k(n)}=1$ the iteration counter $k$ is incremented ($n=n+1$), and the return to the Step 2 of the scheme is carried out. If $h^{k(n)}=0$ the  iterations are completed, and the estimate $e^{k(n)}$  is taken as the (approximate) problem solution.}
\end{description}
\note {In the proposed scheme it is assumed implicitly, that the numbers   of the processors involved into the execution of a particular step of the algorithm  are known a priori for all iterations. At the same time, the scheme can be easily supplemented with a processor number selection rule in dependence on the current search state, if to introduce a functional $P_n(\Phi,\omega_{k(n-1)})$  such that}
\begin{displaymath}
p(n)= P_n(\Phi,\omega_{k(n-1)}).
\end{displaymath}

Then, Step 4 of the computational scheme can be written as follows. 
\begin{description}[\textbf{Step 4}]
\item[\textbf{Step 4}]{The number of processors required for the execution of the  $(n+1)$-th iteration 
\begin{displaymath}
p(n+1)= P_{n+1}(\Phi,\omega_{k(n)}).
\end{displaymath}
is determined and $p(n+1)$  trial coordinates of the current iteration are calculated according to (\ref{eq:24}).}
\end{description}

Let us consider again the method of scanning  a function $\varphi(y)$ in the nodes of a regular grid to illustrate a parallel algorithm. Let us assume for simplicity that the number of processors $p$ is the same for all iterations, that the number of nodes $q>p$  and is divisible by $p$. Then, the decision rules $G(\bullet )$  in the expressions (\ref{eq:20}) and (\ref{eq:24})  obviously take the form 
\begin{displaymath}
G(\bullet )=a+(i-1)\frac{b-a}{q},\ 1\leq i\leq k(n)+1,
\end{displaymath}
and the estimate of extremum and the condition of termination remain the same to those in the sequential prototype.

The synchronous scheme considered here has a drawback that the processors completed their trails within current iteration earlier than other ones, stay idle waiting for the completion of the iteration. Let us consider the computational scheme of an \textit{asynchronous} parallel optimization algorithm, which is free from this disadvantage. In the asynchronous algorithm the iterations become distributed in time. Therefore, it is convenient to deal with the number of trials $k$. In this connection, let us introduce a set of completed trials  $w(k)=\{w_1,w_2,\ldots,w_k\}$  and a set  of coordinates $v(k)=\{v_1,v_2,\ldots,v_{p(k)}\}$ , at which new trials are executed after completing the preceding $k$ trials. Here $p(k)$ is the number of processors executing the current trials, and $w_j=y^{i_j},\ 1\leq j\leq k,\ v_s=y^{i_s},\ 1\leq s\leq p(k), 1\leq i_j,i_s \leq k+p(k)$ . It is obvious, that $w(k)\cup v(k)=\{y^1,\ldots,y^k,\ldots,y^{k+p(k)}\}$ .

Let us assume for simplicity, that we have  $p\geq 2$  processors at our disposal for execution of the algorithm, and this number remains constant in the course of optimization. A variant of the computational scheme with a different number of processors for different search steps can be obtained easily by generalization of the proposed asynchronous scheme consisting in the following.
\paragraph{Asynchronous parallel algorithm scheme}
\begin{description}[\textbf{Step 1}]
\item[\textbf{Step 1}]{First $p$ points of initial trials are selected according to the decision rules $G_i(\bullet ),\ 1\leq i\leq p$ :  
\begin{equation}
\label{eq:26}
\begin{gathered}
y^1=G_1(\Phi\in Q), \\
y^i=G_i(\Phi,y^1,\ldots,y^{i-1)}\in Q,\ 2\leq i\leq p,
\end{gathered}
\end{equation}
the vector $v(0)$  with the components $v_i=y^i,\;1\leq i\leq p$, is formed, and the number of trials $k=0$ is set.}
\item[\textbf{Step 2}]{Wait for the completion of trials by one or several processors. Assume that $1\leq \pi=\pi(k)\leq p$  processors have completed the computation of the function values $z^{i_j}=\varphi(y^{i_j})$  at the points $y^{i_1},\ldots,y^{i_\pi}, y^{i_j}\in v(k),\ 1\leq j\leq \pi$. Then, assign $k=k+\pi$, form the set $w(k)=w(k-\pi)\bigcup \{y^{i_1},\ldots,y^{i_\pi}\}$  and the set $v(k)$ removing from the set $v(k-\pi)$  points $y^{i_1},\ldots,y^{i_\pi}$ . Afterwards there is the following search information on the function $\varphi(y)$ \textit{a posteriori}:
\begin{equation}
\label{eq:27}
\omega_k=\{(y^{i_1},z^{i_1}),(y^{i_2},z^{i_2}),\ldots ,(y^{i_k},z^{i_k}\}.
\end{equation}
where  $y^{i_s}\in w(k),z^{i_s}=\varphi(y^{i_s}),\;1\leq s\leq k$.
}
\item[\textbf{Step 3}]{The current estimate (approximate solution) of the extremum is determined as  
\begin{equation}
\label{eq:28}
e^k=E_k(\Phi,\omega_k).
\end{equation}
}
\item[\textbf{Step 4}]{
\begin{equation}
\label{eq:29}
\begin{gathered}
y^{k+1}=G_{k+1}(\Phi,\omega_{k}) \\
y^i=G_i(\Phi,\omega_k,y^{k+1},\ldots,y^{k+i-1}),\;k+2\leq i\leq k+\pi,
\end{gathered}
\end{equation}
are computed and added to the set $v(k)$.} 
\item[\textbf{Step 5}]{The number 
\begin{equation}
\label{eq:30}
h^k=H_k(\Phi,\omega_k,y^{k+1},\ldots,y^{k+\pi})\in \{0,1\}
\end{equation}
taking one of two possible values: 0 or 1 is computed. If $h^k=1$, switching to Step 2 of the current scheme is carried out. If $h^k=0$, the computations should be completed and the current estimate $e^k$  is taken as the problem solution.}
\end{description}

Some examples of both the synchronous and asynchronous optimization algorithms are presented in the next chapters.
\subsection {Convergence and estimates of extremum}
\label {subsec:1.2.3}
Thus, when solving a minimization problem for a function  $\varphi(y)$ any search method generates a sequence of trial coordinates (or just the trial sequence) $\{y^k\}=y^1,y^2,\ldots,y^k,\ldots$  (where $y^k$   is the coordinate of the $k$-th trial) as well as a sequence of trial results $\{z^k\}=z^1,z^2,\ldots,z^k,\ldots$ . Remind, that as trial results we consider the objective function values, i.e., $z^i=\varphi(y^i),\;i=1,2,\ldots$ .

The behavior of the method is determined via the properties of sequences  $\{y^k\}$ and  $\{z^k\}$. Therefore, the investigation of the search method could be carried out by means of a study of trial sequences which it generates.

Thereby, let us ask the question: what requirements should be imposed to trial sequences of a numerical optimization method? Of course, the main requirement is that execution of the trials at the points $\{y^k\}$  with the trial results $\{z^k\}$  could provide a solution of the problem, i.e., finding the solution corresponding to the selected problem statement. At the same time, since a computer can execute a finite number of trials only, it is desirable to obtain the exact solution after constructing a finite sequence $\{y^k\}$. Therefore,  an asymptotically exact assessment is interesting often, and an infinite trial sequence is considered (in the introduced  optimization method models in (\ref{eq:17}) , (\ref{eq:25}), (\ref{eq:30}) $H_k=1$  for any $k\geq1$, i.e., in fact, the termination criterion is absent). This sequence is required to converge to the exact problem solution. Since in the problem statements $B-D$ the desired solution may include several minimum points, we will understand the convergence of a method in the scope of the following definition.
\begin{definition} 
\label{def:1_4}
The trial sequence $\{y^k\}$  converges to the solution of the problem (\ref{eq:6}) defined by the corresponding problem statement, if:
\begin{enumerate}
\item{it includes a subsequence  $\{\bar{y}^k\}$, for which $\lim_{k \to \infty}\bar{y}^k=\varphi^*$;}
\item{in the case when the solution includes one or several minimum points, for each such point there exists a subsequence of the sequence $\{y^k\}$  converging to it.}
\end{enumerate}
\end{definition}
Trial sequence converging to the exact solution of the problem statement $Z \in \{A,B,C,D,E\}$ is called the minimizing sequence for the problem statement $Z$ or $Z$-minimizing sequence. The term «minimizing sequence» was introduced in \cite{1_Vasiliev} and corresponds to the concept of $A$-minimizing sequence.

Now we would like to pay the reader’s attention to the following important aspect. The property of convergence in the theory of the extremum search methods attracts significant attention, since the asymptotics provides a potential possibility to obtain the exact solution with any predefined accuracy after a finite number of trials. However, this possibility itself is not sufficient for the practical implementation of the methods. In addition, it is necessary to know how to determine a proximity measure of the obtained approximate solution to the exact one, i.e., how to estimate the error of the problem solution using the finite number of trials.

Let us consider the following simple example. Let the class $\Phi$ consist of the continuous one-dimensional functions, defined over the interval $Q=[a,b]$, i.e., it is known a priori that the minimized function $\varphi(y)$ is continuous in the domain $Q$. Assume that the values of the function $\varphi(y)$ are calculated in a finite number of points   What is it possible to say about the coordinate of the global minimum after that? Whatever points $y^1,y^2,\ldots,y^k$  and values $z^1,z^2,\ldots,z^k$ , were obtained after $k$  trials, for any point $y^*\in Q (y^*\notin \{y^1,y^2,\ldots,y^k\})$   a continuous function passing through the points $(y^i,z^i),1\leq i\leq k$,  i.e., belonging to the class $\Phi(\omega_k$ from (\ref{eq:14}) and providing the global minimum at the point $y^*$  with any predefined value
\begin{displaymath}
\varphi^*<\min_{1\leq i\leq k} z^i 
\end{displaymath}
can be built always. For example, an interpolation polynomial of the $k$-th power, passing through the points  $(y^i,z^i),1\leq i\leq k,$ and the point $(y^*,\varphi^*)$, could be taken as such function.

All the above means that no conclusions on the position of the global minimum point can be drawn after any  finite number of trials. Similarly, the only conclusion on the value $\varphi^*$  of the global minimum could be as follows:
\begin{displaymath}
\varphi^*\leq \varphi_k^*, 
\end{displaymath}
where $\varphi_k^*$  is determined by (\ref{eq:19}). Therefore, it is impossible to estimate the value 
\begin{displaymath}
\epsilon_k=\left|\varphi^*-\varphi_k^*\right|, 
\end{displaymath}
i.e., the accuracy of the program solution.

The possibility to obtain the estimates of the extremum based on a finite number of trials depends on the properties of the function class, which the minimized function belongs to, or, in other words, on the information about the function $\varphi(y)$ known a priori. In fact, only two broad classed of the functions allow building such estimates: the class of one-dimensional unimodal functions and the class of the functions satisfying Lipschitz condition (in the general case of the multidimensional and multiextremal functions).

The availability of the extremum estimates built on the base of a finite (truncated) trial sequence allows (in some cases) formulating and solving the problem of theoretical design of the optimal (in different meanings) optimization methods (see, for example, \cite{1_Kushner, 1_Mockus, 1_StrSergMon2000, 1_Sukharev1971, 1_Sukharev1972, 1_Sukharev1989}) . This interesting problem falls out of the scope of the present paper. However, the availability of the extremum estimates is important for the investigation of the parallelization efficiency of the optimization methods considered in the next chapters.

\section{Multiextremal problems and reduction of complexity}
\label{sec:3}
In multiextremal problems the dimension affects the difficulty of solving such problems significantly. For example, for the class of the functions satisfying Lipschitz condition so called «the curse of dimensionality» consisting in the exponential growth of the computational costs with increasing dimensionality takes place. Namely, if $p$ evaluations of the objective function are required to achieve the solution accuracy $\epsilon$ in the one-dimensional case, in the $N$-dimensional problem $\alpha p^N$  trials are required to achieve the same accuracy. Here the coefficient $\alpha$ depends on the objective function, on the feasible domain, and on the method used.

For particular small classes of the multiextremal problems the computational cost growth rate can be slower than the exponential one. For example, for the separable functions minimized over a hyperinterval $D$ the computational costs grow linearly. This fact demonstrates that increase of the efficiency of optimization methods is possible by a thorough accounting for the a priori known information only. Designing the efficient optimization methods as optimal decision rules based on the minimax or Bayesian approach to the concept of optimal methods of the extremum search within the frameworks of the respective mathematical models can be a form of such accounting. Unfortunately, the problem of building the optimal algorithm for the multidimensional multiextremal functions is  very complicated and connected with solving the problems of the covering theory of the feasible domain (see, for example, \cite{1_Sukharev1971, 1_Sukharev1972, 1_Sukharev1989}).

Another approach to the design of the numerical methods for the analysis of the multidimensional multiextremal problems exploits the idea of \textit{the complexity reduction} where the solution of the initial problem is replaced by solving one or several simpler problems.

One of such reduction schemes is based on the elementary fact that the global extremum is a local one. Hence the clear conclusion follows that in order to find the global optimal solution it is sufficient to find all local minima and to select the lowest of them. In the context of the theoretical approach this construct seems to be a perfect one. However, in practice everything is not so rosy. 
First of all, the following question arises: how to find all local minima? This problem could be solved, if for each local minimum its attraction zone, i.e., such neighborhood of the minimum point, in which the function is unimodal, is known. Then, having positioned the initial search point into this neighborhood, one can find the desired local minimum by some local method. In other words, this scheme implies a subdivision of the search domain into the local minimum attraction zones to be known. However, in practice such information is absent a priori as a rule (even the number of local extrema is unknown usually). Therefore, an additional problem of the starting point selection arises when applying this approach.

The simplest method consists in the selection of the starting points using the Monte-Carlo method scheme (see, for example, the fundamental monograph \cite{1_ZhigZhil}), i.e., randomly according to some distribution in the search domain (usually, the uniform one) or in using the nodes of a regular grid as the starting points.

The convergence to the global extremum in this method is provided by the fact that if to increase the number of the starting points at least one of them should fall into the global extremum attraction zone since these grids (either regular or random) are built to provide a coverage of the whole search domain with increasing density. However, this scheme has an essential drawback. As a matter of fact, several starting points can fall into the attraction zone of a local minimum, i.e., this particular local minimum will be found several times. Different schemes have been proposed to overcome this disadvantage ( see \cite{BetroSchoen, BoendeRinnooy, 1_Zielinski}).

The idea of one of them is as follows. First, $L$ base points are selected within the search domain (usually, distributed more or less uniformly). Then, the objective function values are calculated at these points, i.e., actually a rough estimate of the function behavior is performed. Then, the $l<L$  “best” points, i.e., such ones, where the function values are less than at the rest points (at the same time, these “best” points are not too close to each other) are selected from the base points set. These $l$ points serve as the starting points for the local search. There exists a plenty of variants of this scheme. For example, the set of the base points can be modified during the search process: new points can be added to this set. Also, the method of  the point selection may be modified in the course  of the realization of this scheme taking into account the new information obtained, etc.

There also exist other schemes built on the base of the reduction of a multiextremal problem to solving the local subproblems. The algorithms designed within the framework of this approach have an asymptotic convergence to the global extremum with weak presumptions of the continuity or of some degree of differentiability of the objective function. From the quite strict theoretical point of view, this convergence is provided by such property that any point of the search domain is an accumulation point of the trial sequence   generated by the algorithm and, consequently, for a continuous function
\begin{equation}
\label{eq:31}
\lim_{\tau \to \infty}\min_{1\leq k \leq \tau}=\varphi^*.
\end{equation}
In other words, the trial sequences of these methods are everywhere dense in the search domain, i.e., converge in the sense of Definition 1.4 to all points within this domain and, consequently, to the global minimum point as well.

Generally, this character of convergence is too «prodigal», that does not favor the efficiency of the method of this type. Avoiding this drawback is possible only if enough and rather rich a priori known information on the problem is available (like the data on the attraction zones). Such information is rather rarely available in practice and sometimes could be obtained for simple multiextremal problems with a few number of extrema only. For such problems this approach is quite efficient. However, for the essentially multiextremal problems the application of reduction to the local minimum problems is not successful, as a rule. At the same time, the methods of this class are parallelized rather easily, but the everywhere dense convergence (\ref{eq:31}) can results in a large number of redundant trials.

Another approach based on the idea of dimensionality reduction, i.e., on building the schemes, where solving a multidimensional problem is reduced to solving one or several one-dimensional subproblems is more fruitful as compared with the reduction to the local problems. The following two efficient schemes of this kind are known:
\begin{enumerate}
\item{the nested optimization scheme  of optimization \cite{1_CarrHowe, 1_Evtushenko, 1_GerGriIsr, 1_GerGriGer, 1_GriIsrAIP, 1_GriIsrSerg, 1_GriIsrCEUR, 1_Piyavskij, 1_SergGriJCAA, 1_ShiOlaf, 1_StrMonRus, 1_StrSergMon2000, 1_vanDam},}
\item{the reduction of dimensionality based on the curves filling the space (Peano-type curves) \cite{1_Butz, 1_Goertzel, 1_HimOliPet, 1_LeraSergCNSNS, 1_LeraSergANM, 1_SergStrLeraMonogr, 1_StrMonRus,1_StrSergMon2000}.}
\end{enumerate}
The \textit{nested optimization scheme} of dimensionality reduction replaces solving the problem (\ref{eq:8})-(\ref{eq:10}) with solving a family of recursively nested one-dimensional subproblems, in which every evaluation of the objective function within a one-dimensional subproblem consists in solving a new one-dimensional subproblem of the next level of the recursion except the last one, where the value of the initial multidimensional objective function is calculated.

The second scheme is the dimensionality reduction based on the Peano-type curves. It is based on the well known fundamental fact, according to which the $N$-dimensional hyperparallelepiped (\ref{eq:10}) and the interval $[0,1]$ of the real axis are the equinumerous sets, and the interval $[0,1]$ can be mapped onto the parallelepiped (\ref{eq:10}) , unambiguously and continuously \cite{1_SergStrLeraMonogr, 1_StrSergMon2000}. The representations of this kind are usually called \textit{Peano-type curves} or \textit{evolvents}.

Let $y(x),[\in [0,\:1]$  be a Peano-type curve, and the function  $\varphi(y)$ from (\ref{eq:8}) be continuous. Then because of the continuity of  $\varphi(y)$, $y(x)$ and owing to the relation
\begin{displaymath}
D=\{y(x):x\in [0,\:1]\} 
\end{displaymath}
we have the equality
\begin{displaymath}
\min_{y\in D}\varphi(y)=\min_{x\in [0,\;1]}\varphi(y(x)) 
\end{displaymath}
i.e., solving the multidimensional problem of minimization of $\varphi(y)$ over the hyperparallelepiped $D$ is reduced to the minimization of the one-dimensional function $\varphi(y(x))$ over the interval $[0,\;1]$.

The consideration and justification of the algorithms based on the nested optimization scheme and on the dimensionality reduction schemes using Peano-type curves is the major subject of the next chapters. 
% 
\begin{thebibliography}{99.}
\bibitem{BetroSchoen} Betro, B., Schoen, F.: Sequential stopping rules for the multistart algorithm in global optimisation. Math. Program. \textbf{38}(3), 271-286 (1987)
\bibitem{BoendeRinnooy}	Boender, C.G.E, Rinnooy Kan, A.H.G.: Bayesian stopping rules for multistart global optimization methods. Math. Program. \textbf{37}(1), 59-80 (1987)
\bibitem{1_Butz}	Butz, A.R. Space-Filling Curves and Mathematical Programming.: Inform. Control \textbf{12}, 314-330 (1968)
\bibitem{1_CarrHowe}	Carr, C.R., Howe, C.W. Quantitative Decision Procedures in Management and Economic: Deterministic Theory and Applications. McGraw-Hill, New York (1964)
\bibitem{1_Evtushenko}	Evtushenko, Yu.G.: Numerical Optimization Techniques. Translation Series in Mathematics and Engineering. Optimization Software  Inc., Publication Division, New York (1985)
\bibitem{1_FloudasPardalos} Floudas, C.A., Pardalos, P.M.: State of the Art in Global Optimization. Kluwer Academic Publishers, Dordrecht (1996)
\bibitem{1_GerGriIsr} Gergel, V.,  Grishagin, V., Israfilov, R.: Local tuning in nested scheme of global optimization, Procedia Computer Science \textbf{51}, 865-874 (2015) 
\bibitem{1_GerGriGer}	Gergel, V., Grishagin, V., Gergel, A.: Adaptive nested optimization scheme for multidimensional global search, J. Glob. Opt. \textbf{66}, 35–51 (2016).
\bibitem{1_Goertzel}	Goertzel, B.: Global Optimization with Space-Filling Curves. Appl. Math. Lett. 12, 133-135 (1999)
\bibitem{1_GriIsrAIP}	Grishagin, V.A., Israfilov, R.A.: Global search acceleration in the nested optimization scheme. AIP Conf. Proc. \textbf{1738}, 400010 (2016)
\bibitem{1_GriIsrSerg}	Grishagin, V.,  Israfilov, R., Sergeyev, Y.: Convergence conditions and numerical comparison of global optimization methods based on dimensionality reduction schemes. Appl. Math.  Comput. (2017) doi:10.1016/j.amc.2017.06.036
\bibitem{1_GriIsrCEUR}Grishagin, V.A., Israfilov, R.A. Multidimensional Constrained Global Optimization in Domains with Computable Boundaries. CEUR Workshop Proceedings 1513, 75-84 (2015)
\bibitem{1_HimOliPet}	Hime, A.E., Oliveira Jr., H.A., Petraglia, A.: Global Optimization Using Space-Filling Curves and Measure-Preserving Transformations. Soft Computing in Industrial Applications \textbf{96}, 121-130 (2011)
\bibitem{1_Himmelblau}	Himmelblau, D.M.: Applied Nonlinear Programming. McGraw-Hill, New York (1972)
\bibitem{1_HorstPardalos}	Horst, R., Pardalos, P.M.: Handbook of Global Optimization. Kluwer Academic Publishers, Dordrecht (1995)
\bibitem{1_HorstTuy}	Horst, R., Tuy, H.: Global Optimization:Deterministic Approaches. Springer-Verlag, Berlin (1990) (2nd edn., 1993; 3rd edn., 1996)
\bibitem{1_Kushner}	Kushner, H.J.: A new method of locating the maximum point of an arbitrary multipeak curve in the presence of noise. Transactions of ASME, Ser. D. Journal of Basic Engineering \textbf{86}, 97-106 (1964)
\bibitem{1_LeraSergCNSNS}	Lera, D., Sergeyev, Y.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and H{\"o}lder constants. Communications in Nonlinear Science and Numerical Simulation \textbf{23}, 328-342 (2015)
\bibitem{1_LeraSergANM}	Lera, D., Sergeyev, Y.D.: Lipschitz and H{\"o}lder global optimization using space-filling curves. Applied Numerical Mathematics \textbf{60}, 115-129 (2010) 
\bibitem{1_McCormick}	McCormick, G.P.: Nonlinear Programming: Theory, Algorithms and Applications. John Wiley and Sons, New York (1988)
\bibitem{1_Mockus}	Mockus, J.: Bayesian Approach to Global Optimization. Kluwer Academic Publishers, Dordrecht (1988)  
\bibitem{1_Pinter}	Pint{\'e}r, J.D.: Global Optimization in Action.: Kluwer Academic Publishers, Dordrecht (1996) 
\bibitem{1_Piyavskij}	Piyavskij, S.A.: An Algorithm for Finding the Absolute Extremum of a Function. USSR Comput. Math. Math. Phys. \textbf{12}(4), 57-67 (1972)
\bibitem{1_SergGriJCAA}	Sergeyev, Y.D., Grishagin, V.A.: Parallel Asynchronous Global Search and the Nested Optimization Scheme. J. Comp. Analysis Appl. \textbf{3}, 123-145 (2001)
\bibitem{1_SergKvasMonogr}	Sergeyev, Y.D., Kvasov, D.E.: Deterministic Global Optimization: Diagonal Approach Briefly. Springer-Verlag, New York (2017)
\bibitem{1_SergStrLeraMonogr}	Sergeyev, Y.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting Space-Filling Curves. Springer (2013)
\bibitem{1_ShiOlaf}   Shi, L., {\'O}lafsson, S.: Nested Partitions Method for Global Optimization. Operations Research \textbf{48}, 390-407 (2000)
\bibitem{1_StrMonRus}	Strongin, R.G.: Numerical Methods in Multiextremal Problems (Information-Statistical Algorithms). Nauka, Moscow (1978). In Russian
\bibitem{1_StrSergMon2000} Strongin, R.G., Sergeyev, Y.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000) 
\bibitem{1_Sukharev1971}	Sukharev, A.G.: Optimal strategies of the search for an extremum. USSR Comput. Math.Math. Phys. \textbf{11}(4), 119–137 (1971)
\bibitem{1_Sukharev1972}	Sukharev, A.G.: Best sequential search strategies for finding an extremum. USSR Comput. Math. Math. Phys. \textbf{12}(1), 39–59 (1972)
\bibitem{1_Sukharev1989}	Sukharev, A.G.: Minimax Algorithms in Problems of Numerical Analysis. Nauka, Moscow (1989). In Russian
\bibitem{1_vanDam}	van Dam, E.R., Husslage, B., Hertog, D. One-dimensional Nested Maximin Designs. J. Glob. Opt. \textbf{46}, 287-306 (2010)
\bibitem{1_Vasiliev}	Vasiliev, F.P. Numerical Methods for Solving Extremal Problems. Nauka, Moscow (1988). In Russian
\bibitem{1_ZhigZhil} Zhigljavsky, A.A., Zilinskas, A.: Stochastic Global Optimization. Springer, New York. (2008)
\bibitem{1_Zielinski}	Zieli{\'n}ski, R.: A statistical estimate of the structure of multi-extremal problems. Math. Program. \textbf{21}(1), 348-356 (1981)

\end{thebibliography}
%\input{referenc}
%\end{document}

### 2017-09-01 2017/ParOptBook/appendix.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter4.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter5.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter6.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter7.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter8.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\dedic.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\editor.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\foreword.tex
D:\Common\Barkalov\Публикации\20

--- 2017-09-01 2017/ParOptBook/Introduction.tex
@@
-%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%% appendix.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% sample root file for your "contribution" to a contributed volume
+% sample appendix
 %
@@
 %
-%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\documentclass[graybox]{svmult}
-%
-%% choose options for [] as required from the list
-%% in the Reference Guide
-%
-%\usepackage{mathptmx}       % selects Times Roman as basic font
-%\usepackage{helvet}         % selects Helvetica as sans-serif font
-%\usepackage{courier}        % selects Courier as typewriter font
-%\usepackage{type1cm}        % activate if the above 3 fonts are
-                             % not available on your system
-%
-%\usepackage{makeidx}         % allows index generation
-%\usepackage{graphicx}        % standard LaTeX graphics tool
-%                             % when including figure files
-%\usepackage{multicol}        % used for the two-column index
-%\usepackage[bottom]{footmisc}% places footnotes at page bottom
-%
-%% see the list of further useful packages
-%% in the Reference Guide
-%
-%\makeindex             % used for the subject index
-%                       % please use the style svind.ist with
-%                       % your makeindex program
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\begin{document}
+\chapter{Chapter Heading}
+\label{introA} % Always give a unique label
+% use \chaptermark{}
+% to alter or adjust the chapter heading in the running head
 
-\title{Introduction}
-% Use \titlerunning{Short Title} for an abbreviated version of
-% your contribution title if the original one is too long
-\author{Name of First Author and Name of Second Author}
-% Use \authorrunning{Short Title} for an abbreviated version of
-% your contribution title if the original one is too long
-\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
-\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
-%
-% Use the package "url.sty" to avoid
-% problems with special characters
-% used in your e-mail or web address
-%
-\maketitle
-
-\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
-
-\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+Use the template \emph{appendix.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style appendix of your book in the Springer layout.
 
 \section{Section Heading}
-\label{sec:1}
-Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.
-
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Further on please use the \LaTeX\ automatism for all your cross-references and citations.
-And please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
-
-\section{Section Heading}
-\label{sec:2}
+\label{sec:A1}
 % Always give a unique label
@@
 
-Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
+\subsection{Subsection Heading}
+\label{sec:A2}
+Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
+Further on please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:A1}.
 
-Use the standard \verb|equation| environment to typeset your equations, e.g.
-%
-\begin{equation}
-a \times b = c\;,
-\end{equation}
-%
-however, for multiline equations we recommend to use the \verb|eqnarray|
-environment\footnote{In physics texts please activate the class option
-\texttt{vecphys} to depict your vectors in \textbf{\itshape
-boldface-italic} type - as is customary for a wide range of physical
-subjects}.
+For multiline equations we recommend to use the \verb|eqnarray| environment.
 \begin{eqnarray}
-a \times b = c \nonumber\\
-\vec{a} \cdot \vec{b}=\vec{c}
-\label{eq:01}
+\vec{a}\times\vec{b}=\vec{c} \nonumber\\
+\vec{a}\times\vec{b}=\vec{c}
+\label{eq:A01}
 \end{eqnarray}
-
-\subsection{Subsection Heading}
-\label{subsec:2}
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Further on please use the \LaTeX\ automatism for all your cross-references\index{cross-references} and citations\index{citations} as has already been described in Sect.~\ref{sec:2}.
-
-\begin{quotation}
-Please do not use quotation marks when quoting texts! Simply use the \verb|quotation| environment -- it will automatically render Springer's preferred layout.
-\end{quotation}
 
@@
 Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Further on please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{subsec:2}, see also Fig.~\ref{fig:1}\footnote{If you copy text passages, figures, or tables from other works, you must obtain \textit{permission} from the copyright holder (usually the original publisher).
-Please enclose the signed permission with the manuscript.
-The sources\index{permission to print} must be acknowledged either in the captions, as footnotes or in a separate section of the book.}
+Further on please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:A2}.
 
@@
 %
-\begin{figure}[b]
-\sidecaption
-% Use the relevant command for your figure-insertion program
-% to insert the figure file.
-% For example, with the graphicx style use
-\includegraphics[scale=.65]{figure}
-%
-% If no graphics program available, insert a blank space i.e. use
-%\picplace{5cm}{2cm} % Give the correct figure height and width in cm
-%
-\caption{If the width of the figure is less than 7.8 cm use the \texttt{sidecapion} command to flush the caption on the left side of the page.
-If the figure is positioned at the top of the page, align the sidecaption with the top of the figure -- to achieve this you simply need to use the optional argument \texttt{[t]} with the \texttt{sidecaption} command}
-\label{fig:1}       % Give a unique label
-\end{figure}
+%\begin{figure}[t]
+%\sidecaption[t]
+%% Use the relevant command for your figure-insertion program
+%% to insert the figure file.
+%% For example, with the graphicx style use
+%\includegraphics[scale=.65]{figure}
+%%
+%% If no graphics program available, insert a blank space i.e. use
+%%\picplace{5cm}{2cm} % Give the correct figure height and width in cm
+%%
+%\caption{Please write your figure caption here}
+%\label{fig:A1}       % Give a unique label
+%\end{figure}
 
-\paragraph{Paragraph Heading} %
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Further on please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.
-
-Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
-
-For typesetting numbered lists we recommend to use the \verb|enumerate| environment -- it will automatically render Springer's preferred layout.
-
-\begin{enumerate}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\begin{enumerate}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\end{enumerate}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\end{enumerate}
-
-\subparagraph{Subparagraph Heading} In order to avoid simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}, see also Fig.~\ref{fig:2}.
-
-For unnumbered list we recommend to use the \verb|itemize| environment -- it will automatically render Springer's preferred layout.
-
-\begin{itemize}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development, cf. Table~\ref{tab:1}.}
-\begin{itemize}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\end{itemize}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\end{itemize}
-
-\begin{figure}[t]
-\sidecaption[t]
-% Use the relevant command for your figure-insertion program
-% to insert the figure file.
-% For example, with the option graphics use
-\includegraphics[scale=.65]{figure}
-%
-% If no graphics program available, insert a blank space i.e. use
-%\picplace{5cm}{2cm} % Give the correct figure height and width in cm
-%
-%\caption{Please write your figure caption here}
-\caption{If the width of the figure is less than 7.8 cm use the \texttt{sidecapion} command to flush the caption on the left side of the page.
-If the figure is positioned at the top of the page, align the sidecaption with the top of the figure -- to achieve this you simply need to use the optional argument \texttt{[t]} with the \texttt{sidecaption} command}
-\label{fig:2}       % Give a unique label
-\end{figure}
-
-\runinhead{Run-in Heading Boldface Version} Use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.
-
-\subruninhead{Run-in Heading Italic Version} Use the \LaTeX\ automatism for all your cross-refer\-ences and citations as has already been described in Sect.~\ref{sec:2}\index{paragraph}.
-% Use the \index{} command to code your index words
-%
 % For tables use
@@
 \caption{Please write your table caption here}
-\label{tab:1}       % Give a unique label
+\label{tab:A1}       % Give a unique label
 %
@@
 Classes & Subclass & Length & Action Mechanism  \\
-\noalign{\smallskip}\svhline\noalign{\smallskip}
+\noalign{\smallskip}\hline\noalign{\smallskip}
 Translation & mRNA$^a$  & 22 (19--25) & Translation repression, mRNA cleavage\\
@@
 %
-\section{Section Heading}
-\label{sec:3}
-% Always give a unique label
-% and use \ref{<label>} for cross-references
-% and \cite{<label>} for bibliographic references
-% use \sectionmark{}
-% to alter or adjust the section heading in the running head
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Further on please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.
-
-Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
-
-If you want to list definitions or the like we recommend to use the Springer-enhanced \verb|description| environment -- it will automatically render Springer's preferred layout.
-
-\begin{description}[Type 1]
-\item[Type 1]{That addresses central themes pertainng to migration, health, and disease.
-In Sect.~\ref{sec:1}, Wilson discusses the role of human migration in infectious disease distributions and patterns.}
-\item[Type 2]{That addresses central themes pertainng to migration, health, and disease.
-In Sect.~\ref{subsec:2}, Wilson discusses the role of human migration in infectious disease distributions and patterns.}
-\end{description}
-
-\subsection{Subsection Heading} %
-In order to avoid simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Use the \LaTeX\ automatism for all your cross-references and citations citations as has already been described in Sect.~\ref{sec:2}.
-
-Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
-
-\begin{svgraybox}
-If you want to emphasize complete paragraphs of texts we recommend to use the newly defined Springer class option \verb|graybox| and the newly defined environment \verb|svgraybox|.
-This will produce a 15 percent screened box 'behind' your text.
-
-If you want to emphasize complete paragraphs of texts we recommend to use the newly defined Springer class option and environment \verb|svgraybox|.
-This will produce a 15 percent screened box 'behind' your text.
-\end{svgraybox}
-
-\subsubsection{Subsubsection Heading}
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Further on please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.
-
-Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
-
-\begin{theorem}
-Theorem text goes here.
-\end{theorem}
-%
-% or
-%
-\begin{definition}
-Definition text goes here.
-\end{definition}
-
-\begin{proof}
-%\smartqed
-Proof text goes here.
-\qed
-\end{proof}
-
-\paragraph{Paragraph Heading} %
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Further on please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.
-
-Note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
-%
-% For built-in environments use
-%
-\begin{theorem}
-Theorem text goes here.
-\end{theorem}
-%
-\begin{definition}
-Definition text goes here.
-\end{definition}
-%
-\begin{proof}
-\smartqed
-Proof text goes here.
-\qed
-\end{proof}
-%
-\begin{acknowledgement}
-If you want to include acknowledgments of assistance and the like at the end of an individual chapter please use the \verb|acknowledgement| environment -- it will automatically render Springer's preferred layout.
-\end{acknowledgement}
-%
-\section*{Appendix}
-\addcontentsline{toc}{section}{Appendix}
-%
-%
-When placed at the end of a chapter or contribution (as opposed to at the end of the book), the numbering of tables, figures, and equations in the appendix section continues on from that in the main text.
-Hence please \textit{do not} use the \verb|appendix| command when writing an appendix at the end of your chapter or contribution.
-If there is only one the appendix is designated ``Appendix'', or ``Appendix 1'', or ``Appendix 2'', etc.
-if there is more than one.
-
-\begin{equation}
-a \times b = c
-\end{equation}
-
-\input{referenc}
-%\end{document}

### 2017-09-01 2017/ParOptBook/Chapter1.tex

@@
 Second, in numerical solving an optimization problem on a computer the exact calculation of the equality $h(y)=0$ is often impossible, therefore, this equality is assumed to be computed  with some error $\delta>0$ , i.e., the inequality $\left|h(y)\right|\leq \delta$ is considered instead of the equation $h(y)=0$.
-Thus, taking into account the above reasons, one can claim that (\ref{eq:8})-(\ref{eq:10}) is the statement of  \textit {the general  problem of the nonlinear programming}.
+Thus, taking into account the above reasons, one can claim that (\ref{eq:8})--(\ref{eq:10}) is the statement of  \textit {the general  problem of the nonlinear programming}.
 
-If $m=0$ , i.e., there are no functional constraints, let us assume $Q=D$. The problem (\ref{eq:8})-(\ref{eq:10}) in this case will be called \textit {the unconstrained optimization problem}.
+If $m=0$ , i.e., there are no functional constraints, let us assume $Q=D$. The problem (\ref{eq:8})--(\ref{eq:10}) in this case will be called \textit {the unconstrained optimization problem}.
 
@@
 \label{eq:12}
-\varphi(y_1,y_2)=-1.5y_1^2exp(1-y_1^2-20.25(y_1-y_2)^2)-
+\varphi(y_1,y_2)=-1.5y_1^2\exp(1-y_1^2-20.25(y_1-y_2)^2)-
 \end{displaymath}
@@
 \label{eq:12}
--(0.5(y_1-1)(y_2-1))^4exp(2-(0.5(y_1-1))^4-(y_2-1)^4)
+-(0.5(y_1-1)(y_2-1))^4\exp(2-(0.5(y_1-1))^4-(y_2-1)^4)
 \end{displaymath}
@@
 \end{displaymath}
+\begin{displaymath}
+g_2(y_1,y_2)=100(1-(y_1-2)^2/1.44-(0.5y_2)^2)\leq 0,
+\end{displaymath}
+\begin{displaymath}
+g_3(y_1,y_2)=10(y_2-1.5-1.5\sin(6.283(y_1-1.75)))\leq 0,
+\end{displaymath}
+as a particular example.
+
 \begin{figure}[t]
@@
 \end{figure}
-\begin{displaymath}
-g_1(y_1,y_2)=100(1-(y_1-2)^2/1.44-(0.5y_2)^2)\leq 0,
-\end{displaymath}
-\begin{displaymath}
-g_1(y_1,y_2)=10(y_2-1.5-1.5sin(6.283(y_1-1.75)))\leq 0,
-\end{displaymath}
-as a particular example.
+
 The set of the points satisfying the first constraint represents a circle with the boundary defined by the equation $g_1(y_1,y_2)=0$.
@@
 The points satisfying the third constraint are below the sinusoid $g_3(y_1,y_2)=0$.
-Thus, the feasible domain is disconnect and consists of three non-convex subdomains (the feasible domain is dark in Fig.\ref{1_fig_1}).
+Thus, the feasible domain is disconnect and consists of three non-convex subdomains (the feasible domain is dark in Fig.~\ref{1_fig_1}).
 The global minimum $\varphi(y_1^*,y_2^*)=-1.489$  is achieved at the point $(y_1^*,y_2^*)$  = (0,942, 0,944).
@@
 \item{the calculation of the derivative should be possible;}
-\item{it is necessary to solve the equation$\varphi'(y)=0$;}
+\item{it is necessary to solve the equation $\varphi'(y)=0$;}
 \item{the information on the derivative discontinuity points or a method for finding these ones are required.}
@@
 \label{eq:19}
-y_k^*=arg\min_{1\leq i\leq k} \varphi(y^i).
+y_k^*=\arg \min_{1\leq i\leq k} \varphi(y^i).
 \end{displaymath}
@@
 \label {subsec:1.2.2}
-The computational scheme (\ref{eq:12})-(\ref{eq:17}) implies  the sequential execution of all the algorithm steps, that can be provided by the single processor computers.
+The computational scheme (\ref{eq:12})--(\ref{eq:17}) implies  the sequential execution of all the algorithm steps, that can be provided by the single processor computers.
 But if a multiprocessor computer is available, a parallelization of the computational scheme and, therefore, an acceleration of the optimization process can be realized.
 The parallelization of optimization methods could be done in several ways.
-It is possible to parallelize, in the first place, the calculation of the objective function, secondly, the implementation of the rules $G_k,E_k,H_k$ , and, finally, the scheme (\ref{eq:12})-(\ref{eq:17}) providing simultaneous execution of several trials.
+It is possible to parallelize, in the first place, the calculation of the objective function, secondly, the implementation of the rules $G_k,E_k,H_k$ , and, finally, the scheme (\ref{eq:12})--(\ref{eq:17}) providing simultaneous execution of several trials.
 However, the parallelization of the objective function describing the optimized object is specific for different objective function models, and as a matter of fact, cannot be related to the general optimization scheme.
@@
 Let us consider the synchronous variant of the parallelization, where the transition to the next iteration is performed after full termination of the current one, i.e., after completing the last trial of the current iteration.
-In this case, it is obvious that $k(n)=p(1)+\ldots p(n),n\geq 1$.
+In this case, it is obvious that $k(n)=p(1)+\ldots + p(n), \; n\geq 1$.
 Let us denote the vector of the trial coordinates at the $n$-th iteration as $u(n)$ , i.e.,
@@
 \end{enumerate}
-The \textit{nested optimization scheme} of dimensionality reduction replaces solving the problem (\ref{eq:8})-(\ref{eq:10}) with solving a family of recursively nested one-dimensional subproblems, in which every evaluation of the objective function within a one-dimensional subproblem consists in solving a new one-dimensional subproblem of the next level of the recursion except the last one, where the value of the initial multidimensional objective function is calculated.
+The \textit{nested optimization scheme} of dimensionality reduction replaces solving the problem (\ref{eq:8})--(\ref{eq:10}) with solving a family of recursively nested one-dimensional subproblems, in which every evaluation of the objective function within a one-dimensional subproblem consists in solving a new one-dimensional subproblem of the next level of the recursion except the last one, where the value of the initial multidimensional objective function is calculated.
 

### 2017-09-01 2017/ParOptBook/Chapter8.tex

@@
 level) was the object of investigations.
-For the mathematical description of the economical processes taking place at the regional level, a dynamic balance normative economic model has been proposed in Dorodnicyn Computing Centre of RAS [7].
+For the mathematical description of the economical processes taking place at the regional level, a dynamic balance normative economic model has been proposed in Dorodnicyn Computing Centre of RAS \cite{8_Olenev1999}.
 This model is a general one.
@@
 As a result, the problem of identification of the mathematical economic model is the global optimization one.
+
+Once the model was identified, it could be applied to verify various possible scenarios of the regional economic development on the base of the experiments with the model.
+The conclusions and predictions obtained using the identified mathematical model, in principle, could be disposed experimentally.
+This enables to correct further the model in order to obtain more justified prognosis based on the model being improved continuously.
+
+The problem of the identification of the multisector model of the regional economy presented here has been solved on the base of the data available for Nizhni Novgorod region, Russia.
+As it has been already mentioned above, the normative models include a large number of unknown parameters (the normatives of the distribution of the products and the normatives of distribution of the financial resources).
+In order to identify these parameters, it was proposed to apply the efficient parallel global optimization methods.
+The application of these algorithms on the supercomputer systems allows increasing the level of complexity of the mathematical models of the regional economy expressed by the number of the independent parameters.
+Carrying out the scenario computations based on the identified models of the regional economy allows more exact prognosis of the economic consequences of various strategic decisions.
+
+\subsection{The description of the model of the regional economy}
+
+The economy of Nizhni Novgorod region can be aggregated into three main sectors with approximately equal power:
+\begin{itemize}
+	\item the branches of the infrastructure, production and distribution of raw materials (agriculture, electric power industry, development, transportation, state administration, education, public health);
+	\item the manufacturing branches (machine building and metal working, chemical and petrochemical industry, fuel industry, ferrous and nonferrous metallurgy, construction materials industry, timber, woodworking, and pulp-and-paper industry, consumer goods industry, food, pharmaceutical, chemical, microbiological industry, flour-and-cereals and feed mill industry, printing industry, scientific complex);
+	\item the service branches (trading, operations with real estate, financial services, other services).
+\end{itemize}
+
+Let us select the following economic agents while designing the model: Regional Government, the producers represented by the three sectors selected above, the banking system, and the households of the region, the external vendors and consumers.
+
+In spite of absence of rich natural resources in Nizhni Novgorod Region its economy is one of the most developed one with respect to the industry among the regions of Russian Federation.
+The processing branches of industry, a powerful military industrial complex, well-developed fundamental scientific complex are the base of the industry in Nizhni Novgorod Region and there are good prospects for the development of the high-tech industries and of the innovative branches.
+
+The three-sector variant of the general balance model with supplies of products, factors of production, and financial resources in taxation and the illegal circulation of goods and finances \cite{8_Olenev2007} has been applied as the base for the design of the mathematical model of the regional economy.
+Remind that the producers in the model of Nizhni Novgorod Region economy are represented by three sectors:
+\begin{itemize}
+	\item the infrastructure complex ($X$),
+	\item the manufacturing branches complex ($Y$),
+	\item the services, mortgage, financial, and trading branches complex ($Z$).
+\end{itemize}
+The producers (the sectors of the regional economy $X$, $Y$, $Z$) use labor, capital, and the intermediate production of the partner sectors.
+The producers deliver their products to the internal markets and to the external ones.
+The households $L$ offer the labor and consume the final products.
+The resellers $T$ redistribute the goods and financial flows.
+The banking system $B$ lends money to the producers with the purpose to receive the banking profit.
+Regional Government $G$ collects the taxes from the producers and the households and manages the budget expenditures.
+Each market is assumed to establish its own prices for each kind of products.
+The variations of the prices are assumed to be inversely proportional to the variation of the supplies of the corresponding products.
+
+In order to account for the illegal circulation of the goods and money, the producers were assumed to divide the manufactured products into the legal part and the illegal one.
+The latter does not undergo the taxation.
+As a result, the producers receive two kinds of income: the ``legal'' and the ``illegal'' one.
+The ``illegal'' money can be laundered and the reserves undergo penalties.
+
+All the consumers’ money is assumed to be legal.
+The consumers’ income is assumed to be distributed between the acquisition of the legal products and the illegal ones from all sectors according to the predefined percentages.
+The production sectors $m = X,\;Y,\;Z$ pay the profit tax $n_1$, the value added tax $n_2$, the excise tax on the total output $n_3^m$ , the unified social tax on the payroll $n_4$, and the exports customs duties $n_5$.
+The households $L$ within the framework of the model pay the import customs duties $n_6$ and the income tax on the wages $n_7$.
+
+The criteria and the parameters of the model have the upper and lower indices.
+The upper indices are used for the agents and the lower ones for the valuables.
+The supplies of each valuable are distributed according to the normative  $a_i^{nm}$, which are the portions of the amount of the valuable $i$ being transferred from the agent $n$ to the agent $m$.
+The distribution of money is also performed according to some normative $b_i^{nm}$, which are the portions of the money resources of the agent $m$ transferred to the agent $n$ for the product $i$.
+The capital coefficients are also defined by the normatives $c_i^m$ , which are the normalized amount of the product $i$ spent for the producing a unit of a fund product for the agent $m$.
+
+The identification of the parameters was performed by the comparison of the output time series of the model variables with the available statistical time series in 2000--2008.
+As a criterion of proximity of the calculated time series to the statistical ones, Theil index was applied, which has been calculated according to the formula
+\[
+U=\frac{\sqrt{\sum{(x_i-y_i)^2}}}{\sqrt{\sum{x_i^2}}+\sqrt{\sum{y_i^2}}}
+\]
+The close the index to zero, the close the series compared.
+
+\subsection{The results of the experiments}
+
+The detailed description of the designed model is beyond the scope of this paper.
+The additional information can be found in \cite{8_Gergel2011}.
+Here we note that the global optimization problem arising in the model identification process has $60$ independent parameters, $86$ nonlinear constraints, and a multiextremal objective function.
+As an illustration, a two-dimensional section of the objective function confirming its essential multiextremality is presented in Fig.~\ref{8_fig_7}.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.8\linewidth]{figures/8_7.png}
+\caption{Two-dimensional section of the objective function}
+\label{8_fig_7}     
+\end{figure}
+
+This problem has been solved on the cluster of Lobachevsky State University of Nizhni Novgorod using the parallel index method with the rotated evolvents (8 processors were employed).
+The estimate of optimum was obtained after $300 000$ iterations, the computation time of obtaining an estimate was $26$ minutes.
+Total computation time was $3$ hours.
+
+Using the identified mathematical model of Nizhni Novgorod Region, the scenario computations reflecting the dynamics of the economic indicators of the region for various economic policies (the pessimistic scenario, the basic, and the optimistic one) have been carried out in Computing Centre of RAS.
+
+\section*{A brief conclusion}
+
+In this chapter, some applied global optimization problems have been considered.
+All of these problems related with a search for optimal parameters for nonlinear relationships, which describe phenomenon or process under consideration.
+The first example is the finding of the optimum profile of a vehicle wheel.
+Solution to this problem allows to increase the maximum possible speed of the vehicle.
+The second one is related with chemical research (finding the unknown kinetic parameters in the mathematical models of the complex chemical reactions).
+The last example belongs to the field of economics.
+The problem of identification of the parameters for the regional economic model has been solved.
+Optimal parameters found with the global search algorithm allow to carry out scenario forecast calculations for the regional economy.
 
@@
 
-8.
-N.N.
-Olenev.
-Model of innovation potential of regional economy // Proceedings of international conference “Economy of depressed areas: problems and prospects of progress of regional economy”.
-Barnaul, Altai State University Press.
-2007.
-pp.178188.
-(In Russian)
+\bibitem{8_Olenev2007}
+Olenev, N.N.: Model of innovation potential of regional economy.
+Proceedings of international conference ``Economy of depressed areas: problems and prospects of progress of regional economy''.
+Altai State University Press.
+178--188 (2007) (In Russian)
 

### 2017-09-05 2017/ParOptBook/Chapter4.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter4.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter5.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter6.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter7.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter8.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\dedic.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\editor.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\foreword.tex
D:\Common\Barkalov\Публикации\20

--- 2017-09-01 2017/ParOptBook/Chapter7.tex
@@
 %\usepackage{type1cm}        % activate if the above 3 fonts are
-                             % not available on your system
+%                           % not available on your system
 %
@@
 
-\title{Software Systems for Global Optimization }
+\title{Global search algorithms for one-dimensional multiextremal optimization problems with nonlinear constraints}
 % Use \titlerunning{Short Title} for an abbreviated version of
@@
 
-\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\abstract{The chapter considers a method for reduction of a constrained problem to an
+unconstrained one, so called \emph{index method}. This method, unlike the classical penalty function
+one, allows avoiding the problem of the parameter selection (like the penalty coefficients for the
+constraints) and also allows solving the problems with so called partially defined constraints,
+which could not be solved by any other method. The serial and parallel algorithms for solving
+univariate problems are considered.}
 
-\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\keywords{multiextremal optimization, nonconvex constraints, index method, parallel
+algorithms}
 
-\section{Section Heading}
-Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.
+\section{Problem statement}
+In the present chapter, the one-dimensional problems of the constrained optimization will
+be considered.
 

### 2017-09-12 2017/PAVT/paper/Paper_ID80.tex

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb,amsmath}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{url}
\urldef{\mailsa}\path|{alfred.hofmann, ursula.barth, ingrid.haas, frank.holzwarth,|
\urldef{\mailsb}\path|anna.kramer, leonie.kunz, christine.reiss, nicole.sator,|
\urldef{\mailsc}\path|erika.siebert-cole, peter.strasser, lncs}@springer.com|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Optimization of Drop Characteristics in a Carrier Cooled Gas Stream
using ANSYS and Globalizer Software Systems on the PNRPU High-Performance
Cluster\thanks{This study was supported by the Russian Science Foundation,
project No. 16-11-10150.}}

% a short form should be given in case it is too long for the running head
\titlerunning{Optimization of Drop Characteristics in a Carrier Cooled Gas Stream}

\author{Stanislav L. Kalyulin (ORCID 0000-0002-5998-0950)$^1$ \and
Evgenya V. Shavrina$^1$ \and
Vladimir Ya. Modorskii (ORCID 0000-0002-4200-3440)$^1$ \and \\
Konstantin A. Barkalov (ORCID 0000-0001-5273-2471)$^2$ \and
Victor P. Gergel (ORCID
0000-0002-4013-2329)$^2$}
%
\authorrunning{S.L. Kalyulin \and E.V. Shavrina \and V.Ya. Modorskii \and K.A. Barkalov \and V.P. Gergel}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{$^1$Perm National Research Polytechnical University, Perm, Russia\\
\email{\{ksl, modorsky\}@pstu.ru}\\
$^2$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
\email{\{konstantin.barkalov, victor.gergel\}@itmm.unn.ru}
}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem''
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract}
We describe in this article the optimization calculations of spray
droplets in a gas injected through a nozzle into a work area, as a part of
a research of icing on model objects in a small-size climatic wind tunnel.
Calculations were performed in a 3D-statement. It is assumed that the drop
has some speed, temperature and diameter as it enters the gas flow, which
has a specified speed and temperature, so that definite temperature limits
are attained when it interacts with a remote obstruction. We determined
the maximum gas flow temperature, which corresponds to the minimum of
cooling energy consumption. The optimization was carried out using the
Globalizer software (Lobachevsky State University of Nizhny Novgorod).
Also, we could solve the integration issue between Globalizer and ANSYS
Workbench 13.0. ANSYS was employed as a tool to calculate optimization
criteria values, whereas Globalizer was used as an optimal parameter
search tool. Calculations were performed on the PNRPU high-performance
cluster (with a peak performance of 24 TFLOPS).

\keywords{small-size climatic wind tunnel, global optimization, multiextremal
functions, parallel algorithms, droplets flying, numerical simulation, gas
flow.}
\end{abstract}

\section{Introduction}

An energy-efficient \cite{Shmakov2016} closed-loop small-sized climatic wind
tunnel (CWT) is being developed at Perm National Research Polytechnical
University (PNRPU) \cite{Modorskii_1_2016} with a view to continue scientific
investigations into the field of aircraft aerodynamics. This installation which
will provide an opportunity to conduct aerodynamic experiments with simulated
in-flight icing conditions at flow Mach numbers as high as $M=0.8$ and
stagnation temperatures up to 30$^\circ$C \cite{Kalyulin2016}.

As it is known, tests in wind tunnels can help to solve the following problems
\cite{Afanasiev1994}:

\begin{enumerate}
\item Investigation of the influence of the gas-streamlined object shape on
    the object aerodynamic characteristics depending on flow velocity and
    the body position in space.

\item Investigation of air machines: gas turbines, compressors, propellers,
    windmills, fans, etc.

\item Investigation of engine characteristics (piston turbojet, ramjet,
    etc.).

\item Investigation of aircraft flight dynamics.

\item Investigation of the influence of aerodynamic forces on the
    elasticity of aircraft structures (e.g., research into aircraft wing
    flutter).

\item Physical investigations related to air flow under different
    conditions (research into the boundary layer, supersonic flows, spatial
    streams, etc.).

\item Methodical and scientific investigations related to the creation of
    wind tunnels as physical facilities, test method development in tunnels
    and processing of the results obtained.
\end{enumerate}

The article discusses the design of an energy-efficient closed-loop small-sized
CWT for full-scale icing simulation. The modeling of this process requires
structural elements such as air-cooling system, injectors for water supply to
the tunnel, a device to heat them, and a dehumidifying system (described in the
patent \cite{Goryachev2007}).

A distinctive feature of the design is its high energy efficiency, which will
allow for testing under icing conditions at a much lower cost than already
existing large testing facilities. This economy is achieved through its ability
to consume only the energy required in operating mode to maintain a
predetermined level of air flow and by reducing the characteristic dimensions
of the CWT \cite{Kalyulin2016}.

According to studies by Russian scientists \cite{Klemenkov2008}, the
contour of the CWT must be open, so that moisture does not accumulate. The
working part is enclosed as its length is enough to equalize the drop
temperature to air temperature. Owing to the fact that large CWTs are
energy-intensive installations, small-size models are becoming more
popular. However, scale model sizes less than $1:8$ lead to distortion of
the icing shape, which is unacceptable. Thus, there is a limitation on the
minimum diameter of the CWT.

Given the large time and material costs that are necessary for the
preparation and execution of physical experiments in an energy-efficient
closed-loop small-sized CWT, numerical simulation of experiments has
become the most popular method for studying icing processes. We propose
the joint use of numerical and physical experiments.

Ice formation processes during field experiments (spraying water with
subsequent freezing) determine three tasks that must be accomplished for the
numerical simulation of processes occurring in a CWT:
\begin{enumerate}
\item Simulation of cooling gas flowing in the CWT.

\item Simulation of drop disintegration to determine correctly the size of
    ice particles.

\item Simulation of the ``water-ice'' phase transition.
\end{enumerate}

ANSYS CFX numerical experiments concerning spraying of water particles are
possible with the use of different mathematical models of various degrees of
accuracy. In most cases, however, significant computational resources are
required.

A modeling of icing was performed in \cite{Alekseenko2013}, but the method
suggested by the authors does not allow to model the characteristics of
the surface roughness, nor complex-shaped ice build-ups (such as the
unevenness of the ``lobster tail'') in icing areas at small slip angles of
the drops. According to \cite{Prihodko2014}, small build-ups of this kind
have a more significant impact than the ``horn-shaped'' ones. Also, that
paper discusses a technique that allows to estimate the probability of
such ice build-ups. It is not possible, however, to obtain sufficiently
accurate data on the nature of the build-ups in such processes as flow
separation from a wing.

Icing simulation in ANSYS FENSAP-ICE has not gained much popularity due to
requirements to conclude additional license agreements and an insufficient
number of validation studies. Nevertheless, a solution example of the
problem of icing in ANSYS FENSAP-ICE in conjunction with ANSYS CFX was
presented by researchers from Louisiana in \cite{Hannat2014}. ANSYS CFX
and ANSYS FLUENT gas dynamics packages are usually used in this regard in
research practice.

As a rule, information on temperature, humidity, speed and other gas
dynamics parameters are used to predict icing,  since a direct modeling of
the ``water-ice'' phase transition is not possible in these packages.
Besides, investigations on gas dynamics problems performed abroad with the
use of the ANSYS CFX and ANSYS FLUENT packages relied on subsequent
accurate gas dynamics parameters to solve the problem of icing by
semi-empirical or analytical methods \cite{Villalpando2016}.

The statement of the problem of numerical simulation depends on the type
of icing, which can be divided into four categories
\cite{Klemenkov2008,Lynch2001,Bragg2005}:

\begin{enumerate}
\item The roughness of the ice surface surpasses the height of the
    local boundary layer, affecting not only the transition of the
    boundary layer to a turbulent state, but also its separation
    downstream.

\item The icing is characterized by grooves along the stream. In this
    case, the area of flow separation primarily depends on the attack
    angle. In combination with its roughness, this type of icing
    mainly affects the increase in resistance.

\item In some cases, the droplets spreading over the surface
    crystallize immediately. Then the ice begins to grow in a vicinity
    of the critical point adopting the shape of a ``horn''. In other
    cases, crystallization is delayed; a film of water creeps over the
    surface, and ice forms along two areas on the side of the critical
    point, forming two ``horns''. The icy ``horns'' create a large
    separated region which affects aerodynamics. This is accompanied
    by the separation of flow depending on their shape, position, and
    angle of attack. It significantly reduces the aircraft’s
    load-bearing properties.

\item When a thermal anti-icing system is working, a film of water
    moves across the surface to the outer boundary of the heated
    region and then freezes in the form of a so-called barrier ice. An
    obstacle is formed which leads to the appearance of a large
    separated region in front and behind. This significantly affects
    the aerodynamic characteristics depending on the position and
    geometry of the obstacle \cite{Bragg2005}.
\end{enumerate}

The nature of these phenomena and the conditions producing a particular
type of icing have not been thoroughly studied until now.

Furthermore, the estimation of the surface tension of small droplets with
a typical diameter of 20~$\mu$m have shown that the excess pressure, which
could affect the freezing point of water and the crystallization process,
is so small that it practically does not affect the freezing temperature
of water \cite{Klemenkov2008}. The surface tension only has an effect on
the spreading of drops when they hit the surface and the disintegration of
the liquid into drops at being expelled from the injectors
\cite{Modorskii_2_2016}. These processes are poorly understood and will be
the subject of forthcoming experimental studies in CWT.

In this regard, mathematical modeling of icing at the moment is not
perfect. Therefore, physical experiments cannot be completely excluded
from consideration. Thus, to maximize at present the effectiveness of
research into icing processes, both numerical and physical experiments are
equally needed.

Investigations are frequently limited to finding one or several solutions
that satisfy the requirements of technical specifications. Searching for
an optimal solution is usually not possible, since it requires  large
computational resources.

Therefore, the numerical search for optimal solutions in an acceptable
time frame is one of the most important tasks
\cite{Gaynutdinova2015,Modorskii_3_2016}. It increases the quality of
decisions and enhances the benefits of numerical calculations, enabling
the identification of new dependencies.

\section{Gas dynamics problem}

\subsection{A physical model of the gas dynamics problem}

In the physical statement of the problem, it is assumed that a drop with a
given velocity, temperature and diameter falls into a gas dynamic flow
having some speed and temperature, so that, upon contact with a distant
obstacle, certain temperature limits are attained \cite{Modorskii2002}.
The distance to the obstacle is equal to 2~m. The maximum temperature of
the carrying gas dynamic flow is determined, and it corresponds to the
minimum energy consumption of the energy-efficient closed-loop small-sized
CWT when it cools.

The calculation is carried out in a 3D dynamic statement. The internal CWT
volume is modeled, and the two-phase environment, gravity and interaction
with the cooling gas dynamic flow are taken into account.

\subsection{Mathematical model of the gas dynamics problem}

We applied the finite volume method for the numerical simulation of the
drop flight from the site of injection through the nozzle to the working
area of the CWT. This implies that the field of the gas dynamic flow is
approximated by a finite number of calculation points.

A mathematical model is developed in accordance with the adopted physical
model. This model is based on the mass, momentum and energy conservation
laws, and is closed with the state equations of an ideal compressible gas
and turbulence, as well as with initial and boundary conditions. The
mathematical model of the gas dynamics problem includes the following
relations \cite{Kozlova2010}:
\begin{itemize}
\item[$\bullet$] The equation of mass conservation for the gas:
\begin{equation}
\frac{\partial}{\partial t}(\rho_{\text{air}})+\bar{\nabla}\cdot(\rho_{\text{air}} V_{\text{air}})=Q^{\text{drop}}_{\text{mass}},
\end{equation}
where $\rho_{\text{air}}$ is the air density, $V_{\text{air}}$ is the
air velocity vector, and $Q^{\text{drop}}_{\text{mass}}$ is a source
term expressing the increase in weight caused by evaporation of water
from the drop surface.

\item[$\bullet$] The equation of momentum conservation for the gas:
\begin{equation}
\frac{\partial}{\partial t}(\rho_{\text{air}} V_{\text{air}})+(\rho_{\text{air}} V_{\text{air}}\cdot \bar{\nabla})\cdot V_{\text{air}}=-\bar{\nabla}P+\bar{\nabla}\tau_{\text{air}}+\rho_{\text{air}}g+Q^{\text{drop}}_{\text{force}},
\end{equation}
where $P$ is pressure, $g$ is the gravity vector, $\tau_{\text{air}}$
is the shear stress at the wall, and $Q^{\text{drop}}_{\text{force}}$
is a source term that expresses the force with which the drop acts on
the air.

\item[$\bullet$] The equation of energy conservation for the gas:
\begin{equation}
\frac{\partial}{\partial t}(\rho_{\text{air}} H_{\text{air}})+\bar{\nabla}\cdot(\rho_{\text{air}} V_{\text{air}} H_{\text{air}})=\frac{\partial P}{\partial t}\cdot\bar{\nabla}\cdot \left(\left(\frac{\lambda_{\text{air}}}{c_{p_{\text{air}}}}+\mu_t\right)\bar{\nabla}H_{\text{air}}\right)+Q^{\text{drop}}_{\text{energy}},
\end{equation}
where $H_{\text{air}}$ is the total enthalpy of the air,
$\lambda_{\text{air}}$ is the coefficient of thermal conductivity of
the air, $c_{p_{\text{air}}}$ is the air heat capacity, $\mu_t$ is
turbulent dynamic viscosity, $Q^{\text{drop}}_{\text{energy}}$ is a
source term expressing energy transfer between the phases.

\item[$\bullet$] The equation of mass conservation for the steam:
\begin{equation}
\frac{\partial}{\partial t}(\rho_{\text{air}} Y_{\text{steam}})+\bar{\nabla}\cdot(\rho_{\text{air}} V_{\text{air}} Y_{\text{steam}})=\bar{\nabla}\cdot\left( \left( \rho_{\text{air}} D + \frac{\mu_t}{S_{c_t}}\right)\bar{\nabla} Y_{\text{steam}}\right)+Q^{\text{drop}}_{\text{mass}},
\end{equation}
where $Y_{\text{steam}}=1-Y_{\text{air}}$ is steam mass concentration,
$Y_{\text{air}}$ is air mass concentration, $D$ is the diffusion
coefficient, and $S_{c_t}$ is the Schmidt turbulent number
($S_{c_t}=1$).

\item[$\bullet$] The equation of state:
\begin{equation}
P=\frac{\rho_{\text{air}}R_0 T_{\text{air}}}{M},
\end{equation}
where $R_0=8314.41~\text{J}/\text{kmol}\cdot \text{K}$ is the
universal gas constant, $M$ is the molecular weight, and
$T_{\text{air}}$ is the air temperature.

\item[$\bullet$] The equation of the drop motion dynamics:
\begin{equation}
\frac{\partial V_{\text{drop}}}{\partial t}=\frac{\pi d^2_{\text{drop}}}{8 m_{\text{drop}}}C_D\rho_{\text{drop}}|V_{\text{air}}-V_{\text{drop}}|(V_{\text{air}}-V_{\text{drop}})+g\left(1-\frac{\rho_{\text{air}}}{\rho_{\text{drop}}}\right),
\end{equation}
where $d_{\text{drop}}$ is the drop diameter, $m_{\text{drop}}$ is the
drop weight, $\rho_{\text{drop}}$ is the drop density,
$V_{\text{drop}}$ is drop velocity, and $C_D$ is the coefficient of
resistance.

\item[$\bullet$] The equation of mass conservation for the drop:
\begin{equation}
\frac{\partial m_{\text{drop}}}{\partial t}=-m^*\pi d^2_{\text{drop}},
\end{equation}
where $m^*$ is the steam injection parameter per drop surface unit.

\item[$\bullet$] The equation of energy conservation for the drop:
\begin{equation}
\frac{\partial T_{\text{drop}}}{\partial t}=\left(\left[\mathrm{Nu}\frac{\lambda_{\text{air}}}{d_{\text{drop}}}(T_{\text{air}}-T_{\text{drop}})\right]-m^* q(T_{\text{drop}})\right)\frac{6}{d_{\text{drop}}\rho_{\text{drop}}c_{p_{\text{drop}}}},
\end{equation}
where $c_{p_{\text{drop}}}$ is the drop heat capacity,
$T_{\text{drop}}$ is the drop temperature, and $\mathrm{Nu}$ is the
Nusselt number.

\item[$\bullet$] The equation of turbulent energy:
\begin{equation}
\frac{\partial}{\partial t}(\rho_{\text{air}}K)+\bar{\nabla}\cdot(\rho_{\text{air}} V_{\text{air}}K)=\bar{\nabla}\cdot\left(\left(\mu+\frac{\mu_t}{\sigma}\right)\bar{\nabla}K\right)+\mu_t G-\rho_{\text{air}}\varepsilon,
\end{equation}
where $\sigma$ is a constant, $G$ determines the speed of turbulent
energy generation, $\varepsilon$ is the rate of turbulent energy
dissipation, and $K$ is turbulent energy.

Given that the computational domain has a cylindrical geometry, the
gas dynamic flow has a high speed, and the length is small and equals
the distance from the injection of droplets through the nozzles to the
barrier (experimental model), it is advisable to use the
$K-\varepsilon$ turbulence model, which is applicable when the
influence of inertial forces is large compared to viscosity forces.

\item[$\bullet$] The equation of turbulent energy dissipation rate:
\begin{equation}
\frac{\partial}{\partial t}(\rho_{\text{air}}\varepsilon)+\bar{\nabla}\cdot(\rho_{\text{air}} V_{\text{air}}\varepsilon)=\bar{\nabla}\cdot\left(\left(\mu+\frac{\mu_t}{\sigma_\varepsilon}\right)\bar{\nabla}\varepsilon\right)+C_1 \frac{\varepsilon}{K}\mu_t G-C_2 f_1 \rho_{\text{air}}\frac{\varepsilon^2}{K},
\end{equation}
where $\sigma_\varepsilon$, $C_1$ and $C_2$ are  constants.
\end{itemize}

\subsection{Parameters and criteria for optimization of gas dynamics calculation}

The problem is solved in ANSYS CFX, as the initial and boundary conditions
are given. We studied the solution convergence. Next, input and output
parameters were parameterized in CFX and transferred to the ANSYS
Workbench, a step that was necessary for the iterative start of ANSYS in
the Globalizer optimization program.

The following ranges of input parameters were set:
\begin{enumerate}
\item $T_{\text{air}}=- 30$ .. $0\,^\circ $C for gas dynamic flow
    temperature;

\item $V_{\text{air}}= $ 10 .. 270~m/s for gas dynamic flow rate;

\item $T_{\text{drop}}|_{l=0} =+5$ .. $+10 \,^\circ$C for the drop
    temperature at the initial moment of contact with the air flow;

\item $V_{\text{drop}}=  10$ .. $270$~m/s for the drop velocity at the
    initial time of contact with the air flow.
\end{enumerate}

The average drop diameter for injecting was taken equal to 20~$\mu$m, and
did not change during computations.

We adopted the following output criteria for the optimization algorithm:
\begin{enumerate}
\item $T_{\text{drop}}|_{l=0} =-0.5$ .. $+0.5\,^\circ$C for the drop
    temperature at the moment when it reaches the obstacle ($L= 2$~m);

\item $T_{\text{air}}\rightarrow \text{max}$ for the maximum initial
    temperature of the gas dynamic flow.
\end{enumerate}

The input parameter $T_{\text{air}}$ is simultaneously an output criterion
that approaches the maximum.

As we know, when integrating the multi-criteria parallel optimization IOSO
PM program complex with ANSYS, it is possible to set the input parameter
as an output optimization criterion only with additional program code
associated with syntactic IOSO PM parameters, i.e. it is not foreseen
within the core function of the optimization complex.

As regards the drop freezing, the air temperature cannot physically be
above $+0.5\,^\circ$C, otherwise the drop will not be able to cool down to
the temperature defined by the output criterion.

The optimization problem was solved with the generalized global
optimization algorithm implemented in Globalizer. To allow for sorting the
results of the Globalizer’s gas dynamics solution in ANSYS CFX, we wrote
the respective functions, and parameterized the total count time
(TotalTime) and the time step (TimeSteps).

The description of both the optimization algorithm and the Globalizer
program system is given in the next section.

\section{Global optimization with non-convex constraints}

\subsection{Problem statement}

Let us consider the $N$-dimensional global optimization problem
\begin{equation}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
\end{equation}
with search domain
\begin{equation}\label{D}
D=\left\{y\in R^N: -2^{-1}\leq y_j \leq 2^{-1}, 1\leq j \leq N \right\}.
%D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.
\end{equation}
This problem statement covers a large class of problems since the hyperinterval
\[%\label{S}
S=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}
\]
can be reduced to the hypercube (\ref{D}) by linear transformation. The
objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and
the left-hand sides $g_i(y), \; 1\leq i \leq m$, of the constraints
satisfy the Lipschitz conditions with constants $L_i, \; 1 \leq i \leq
m+1$, respectively,  and may be multiextremal. It is assumed that the
functions $g_i(y)$ are defined and computable only at the points $y \in D$
satisfying the conditions
\begin{equation}\label{g_k}
g_k(y) \leq 0, \; 1 \leq k < i.
\end{equation}

Employing the continuous single-valued Peano curve $y(x)$
(\textit{evolvent}) that maps the unit interval $[0,1]$ of the $x$-axis
onto the $N$-dimensional domain (\ref{D}), it is possible to find the
minimum in the problem (\ref{problem}) by solving the one-dimensional
problem
\[
\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\},
\]
where, as it follows from(\ref{g_k}), the functions $g_i(y(x))$ are
defined and computable in the domains
\[
Q_1=[0,1], \; Q_{i+1}=\left\{x \in Q_i : g_i(y(x)) \leq 0 \right\}, \; 1 \leq i \leq m.
\]

These conditions allow for the introduction of a classification of the
points $x \in [0,1]$ according to the number $\nu (x)$ of the constraints
computed at this point. The \textit{index} $\nu(x)$ can also be defined by
the conditions
\[
g_i(y(x)) \leq 0, \; 1 \leq i < \nu, \; g_\nu(y(x))>0,
\]
where the last inequality is inessential if $\nu=m+1$.

The considered dimensionality reduction scheme juxtaposes a
multidimensional problem with lipschitzian functions to a one-dimensional
problem where the corresponding functions satisfy the uniform H{\"o}lder
condition (see \cite{Strongin2000}), i.e.
\[
\left|g_i(y(x'))-g_i (y(x''))\right| \leq H_i \left|x'-x'' \right|^{1/N}, \; x',x''\in [0,1], \; 1\leq i \leq m+1.
\]
Here $N$ is the dimensionality of the initial multidimensional problem and
the coefficients $H_i$ are related with the Lipschitz constants $L_i$ of
the initial problem by the inequalities $H_i \leq 2L_i \sqrt{N+3}$.

Thus, a \textit{trial} at a point $x^k \in [0,1]$ executed at the $k$-th
iteration of the algorithm will consist in the following sequence of
operations:
\begin{itemize}
	\item Determine the \textit{image} $y^k=y(x^k)$ in accordance with
the mapping $y(x)$.

	\item Compute the values $g_1(y^k),...,g_\nu(y^k),$ where the
index $\nu \leq m$ is determined by the conditions
	\[
		g_i(y^k )\leq 0, \; 1 \leq i < \nu, \; g_\nu(y^k)>0, \; \nu \leq m.
	\]
\end{itemize}
The occurrence of the first violation of the constraint terminates the
trial. In the case when the point $y^k$ is a feasible one, i.e. when
$y(x^k)\in Q_{m+1}$, the trial includes the computation of the values of
all the functions of the problems and the index is assumed to be
$\nu=m+1$. The pair of values
\[
\nu=\nu(x^k), \; z^k=g_\nu(y(x^k))
\]
is a \textit{trial result}.

\subsection{Generalized global search algorithm}

The rules that determine the work of the index method are the following.

The first trial is executed at an arbitrary internal point $x_1 \in
(0,1)$. The selection of the point $x^{k+1}, \; k \geq 1,$ of any next
trial is determined by the following rules.

Rule 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the
lower indices in ascending order of coordinate values, i.e.
\[
0=x_0<x_1<\dots <x_k<x_{k+1}=1,
\]
and juxtapose to them the values $z_i=g_\nu(y(x_i)), \; \nu=\nu(x_i), \; 1
\leq i \leq k$, computed at these points. The points $x_0=0$ and
$x_{k+1}=1$ are introduced additionally, while the values $z_0$ and
$z_{k+1}$ are not defined.

Rule 2. Classify the indices $i, \; 1 \leq i \leq k$, of the trial points
according to the number of the problem constraints fulfilled at these
points, by constructing the sets
\[
I_\nu =\left\{i:1 \leq i \leq k, \; \nu=\nu(x_i) \right\}, \; 1 \leq \nu \leq m+1,
\]
containing the numbers of all the points $x_i, \; 1 \leq i \leq k$, with
the same values of $\nu$. The end points $x_0=0$ and $x_{k+1}=1$ are
interpreted as the ones having indices equal to zero. An additional set,
$I_0=\left\{0,k+1\right\}$, corresponds to them.

Determine the maximum value of the index:
\[
M=\max\left\{\nu(x_i), \; 1 \leq i \leq k \right \}.
\]

Rule 3. Compute the current lower estimates,
\[
\mu = \max\left\{ \frac{\left|z_i-z_j\right|}{ (x_i - x_j)^{1/N} }, \; i,j \in I_\nu, \; i>j \right\},
\]
for the unknown H{\"o}lder constants $H_\nu$ of the functions $g_\nu(y),1
\leq \nu \leq m+1$. If a set $I_\nu$ contains less than two elements, or
if $\mu_\nu$ is equal to zero, then assume $\mu_\nu=1$.

Rule 4. For all nonempty sets $I_\nu, \; 1 \leq \nu \leq m+1$, compute the
estimates
\[
z_\nu^\ast = \left\{
   \begin{array}{lr}
     -\epsilon_\nu, & \nu < M,\\
     \min\{ g_\nu(y(x_i): i\in I_\nu \}, & \nu = M,
   \end{array}
\right.
 \]
where the nonnegative numbers $(\epsilon_1,...,\epsilon_m)$ are parameters
of the algorithm.

Rule 5. For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute
the \textit{characteristics} $R(i)$ :
\[
R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
\]
\begin{equation}%\label{eq:14}
R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
\end{equation}
\[
R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
\]
where $\Delta_i=(x_i - x_{i-1})^{1/N}$. The values $r_\nu > 1, \; 1 \leq
\nu \leq m+1,$ are parameters of the algorithm. An appropriate selection
of $r_\nu$ allows to consider the product $r_\nu \mu_\nu$ as an estimate
of the H{\"o}lder constants $H_\nu, \; 1 \leq \nu \leq m+1$.

Rule 6. Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
\begin{equation}\label{MaxR}
R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
\end{equation}

Rule 7. Make the next trial at the midpoint of the interval
$(x_{t-1},x_t)$ if the indices of the points $x_{t-1}$ and $x_t$  are not
the same, i.e.
\[
x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
\]
Otherwise, make the trial at the point
\begin{equation}%\label{eq:142}
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
\end{equation}

We may take as termination condition the inequality $\Delta_t \leq
\epsilon$, where $t$ is from (\ref{MaxR}) and $\epsilon>0$ is the
predefined accuracy.

Various modifications of this algorithm and the corresponding theory of
convergence are given in
\cite{Strongin2000,Barkalov2002,Gergel2005,Strongin2013}.

The algorithm considered above is very flexible and allows for an
efficient parallelization for shared memory, for distributed memory, and
for accelerators \cite{Barkalov2010,Barkalov2015,Barkalov2016,Gergel2016}.

\section{Integration of Globalizer and ANSYS Workbench}

This section contains a brief explanation of the Globalizer software
system. The system expands the family of global optimization software that
has been consistently developed at Nizhny Novgorod State University during
the past several years.

The development of the system was based on the generalized global search
algorithm that was described in the previous section. A major advantage of
Globalizer is that the system is designed to solve time-consuming
multiextremal optimization problems. The system efficiently uses modern
high-performance computer systems to obtain the optimal solution within a
reasonable time and at a reasonable cost.

The structural components of the Globalizer are the following blocks.
\begin{itemize}
   \item Block 0 consists of the procedures for computing the function
       values (objective and constraints) for the optimization problem
       that is being solved; this is an external block with respect to
       the Globalizer system.

   \item Blocks 1--3 form the optimization subsystem and solve
       unconstrained (Block 1), constrained (Block 2) and
       multicriteria (Block 3) global optimization problems.

   \item Block 4 is a subsystem for accumulating and processing search
       information.

   \item Block 5 contains the dimensional reduction procedures based
       on evolvents; the optimization blocks solve the reduced
       (one-dimensional) optimization problems. Block 5 provides for
       interaction between the optimization blocks and the initial
       multidimensional optimization problem.

   \item Block 6 is responsible for managing the parallel processes
       while performing the global search (determining the optimal
       configuration of parallel processes, distributing the processes
       between computer elements, balancing computational loads,
       etc.).

   \item Block 7 is a common managing subsystem, which fully controls
       the entire computational process when solving global
       optimization problems.
\end{itemize}
In order to integrate Globalizer with ANSYS Workbench 13.0, we need to
provide the call of an external calculator of function values (i.e. ANSYS)
as Block 0. It is necessary to transfer the input parameter values of the
ANSYS Workbench project, update the project and  receive the output
parameter values. The transfer of the parameters between the ANSYS
Workbench and Globalizer was implemented by means of an IronPython script.

In the conducted experiments, computations were performed on a single
cluster node. The parallelization of computations of the problem function
values relied on ANSYS tools. The computation costs of executing the
optimization algorithm in Globalizer (the choice of the next trial points,
processing the trial results, etc.) were significantly less than those  of
computing the problem function values in ANSYS. Therefore, it was not
necessary to parallelize the operations of the optimization module itself.

It is important to note that the integration of ANSYS and Globalizer
allows for larger scale experiments to be performed with the use of
several cluster nodes as well. In this case, search trials at different
points of the search domain will be performed at all the nodes that are
employed at each iteration of the method. Therefore, the results of the
computations can be processed in Globalizer both synchronously and
asynchronously. In the first case, the system waits until the computations
on all nodes employed are completed. Then the processing of the results is
performed, and the next iteration is started. This regime is preferred at
equal computing time at any point of the search domain. In the second
case, the results obtained at one of the nodes are processed immediately,
and the next trial begins at this node. The asynchronous regime provides
for the full load of the nodes to be utilized in the cases when trial
times are different at different points of the search domain. The
experiments performed earlier on test problems have demonstrated this
approach to be promising \cite{Gergel2015}. Performing experiments of this
kind on the solution of applied problems will be the subject of further
investigations.

\section{Optimization results}

The Globalizer system allowed to construct a field of input parameter
values that satisfy the above-mentioned (Sec. 2.3) restrictions of the gas
dynamics problem (Fig.~\ref{fig1}).

The total time spent on finding the solution of the optimization problem
at the PNRPU high-performance computing complex was 48 hours.

\begin{figure}
\centering
\includegraphics[height=6 cm]{fig1}
\caption {Field of input parameter values that satisfy the output criteria of the
optimization problem: $V_{\text{drop}}$~--- drop velocity, $T_{\text{drop}}|_{l=0}$~---
drop temperature at the time of injection, $V_{\text{air}}$~--- air velocity}
\label{fig1}
\end{figure}

It was assumed that the energy-efficient modes of the CWT are those in
which the gas dynamic flow temperature in the working section approaches
the maximum at the drop temperature $+0.5\,^\circ$C when it reaches the
barrier at a 2~m distance from the injection point:

$T_{\text{air}}\rightarrow \text{max}$;

$T_{\text{drop}}|_{l=2~\text{m}} = 0.5 \,^\circ$C.

However, according to expert estimates, the ranges of permissible drop
speed $V_{\text{drop}}$ and temperature $T_{\text{drop}}|_{l=0}$ at the
time of injection are the following:

$V_{\text{drop}}= $ 5 .. 270~m/s;

$T_{{\text{drop}}|_{l=0}} =+$5 .. $+$10 $^\circ$C.

As the optimization computations show, the values of drop speed and
temperature in the energy-efficient modes of the CWT can vary within the
following ranges (Fig.~\ref{fig1}):

$V_{\text{drop}}= 42.37$ .. 237.63~m/s;

$T_{\text{drop}}|_{l=0} =+5$ .. $+9\,^\circ$C.

An analysis of the results of the computational optimization experiment
(Fig.~\ref{fig1}) showed that the range of drop velocities
$V_{\text{drop}}$ and drop temperatures $T_{\text{drop}}|_{l=0}$
corresponding to these modes of operation is narrower than the recommended
pre-peer evaluations.

These findings identify the speed (for 26.32\%) and temperature (for 20\%)
of the injection modes that produce icing in the work area based at the
specified distance of the experimental model to the injection zone.

Fig.~\ref{fig2} shows that the variation of only two parameters, such as
the speed of the gas dynamic flow ($V_{\text{air}}$) and the drop
temperature at the time of injection through the nozzles
($T_{\text{drop}}|_{l=0}$), allows to achieve the desired value of the
drop temperature ($+0.5\,^\circ$C) upon reaching the barrier (experimental
model) when meeting the output optimization criterion
$T_{\text{air}}\rightarrow+0.5 \,^\circ$C. This may help to achieve
energy-efficient modes in the CWT.

\begin{figure}
\centering
\includegraphics[height=6 cm]{fig2}
\caption {Dependence of $V_{\text{air}}$ (gas dynamic flow rate) and $T_{\text{air}}$
(gas dynamic flow temperature) on $T_{\text{drop}}|_{l=2~\text{m}}$ (drop temperature
when reaching the obstacle ($L= 2$~m)) for different values of $T_{\text{drop}}|_{l=0}$
(drop temperature at time of injection)}
\label{fig2}
\end{figure}

For example, in the given operation mode of the studied experimental
model, by varying only technological parameters such as temperature and
velocity of drops at the initial moment of contact with the gas flow, one
can achieve an energy-efficient simulation of ice formation, except for a
few bottlenecks in the drop velocity modes. Moreover, areas can be
specified by applying a greater number of iterations (ANSYS starts) in the
Globalizer optimization algorithm, which, of course, requires longer
computation times.

\section{Conclusions}

\begin{enumerate}
\item As a result of the solution to the optimization problem, a
    combined region of speed and temperature parameters has been
    detected, as well as the air flow velocity at which the gas
    dynamic flow temperature in the working section approaches the
    maximum. This region is energy efficient, since it allows for
    icing without achieving significant negative airflow temperatures.

\item The possibility of implementing an energy efficient operation
    mode of the small-sized closed-loop CWT has been demonstrated by
    changing only technological parameters such as the initial drop
    temperature and velocity.

\item The integration of the Globalizer optimization system and the
    engineering software package ANSYS has been realized for the first
    time.

\item The Globalizer software package helps to find the best solutions
    at reasonable time and material costs.

\item The application of the Globalizer optimization software system
    made it possible to identify a fairly wide range of parameter
    combinations that allows to maintain the energy-efficient
    operation mode of the small-sized closed-loop CWT with sufficient
    accuracy for engineering practice ($\pm1\,^\circ$C).
\end{enumerate}

\begin{thebibliography}{25}

\bibitem{Modorskii_1_2016}
Modorskii, V.Ya., Shevelev, N.A.: Research of aerohydrodynamic and aeroelastic processes on PNRPU HPC system. In: Vasily Fomin (Ed.) ICMAR 2016, AIP Conference Proceedings, vol. 1770, art. no. 020001 (2016); DOI: 10.1063/1.4963924

\bibitem{Shmakov2016}
Shmakov, A.F., Modorskii, V.Y.: Energy Conservation in Cooling Systems at Metallurgical Plants. Metallurgist 59(9--10), 882--886 (2016); DOI: 10.1007/s11015-016-0188-8

\bibitem{Kalyulin2016}
Kalyulin, S.L., Modorskii, V.Ya., Paduchev, A.P.: Numerical design of the rectifying lattices in a small-sized wind tunnel. In: Vasily Fomin (Ed.) ICMAR 2016, AIP Conference Proceedings, vol. 1770, art. no. 030110 (2016); DOI: 10.1063/1.4964052

\bibitem{Afanasiev1994}
Afanasiev, V.A., Barsukov, V.S., Gofin, M.Ya., Zakharov, Yu.V., Strelchenko, A.N., Shalunov, N.P.: Experimental testing of spacecraft. MAI, Moscow (1994) [in Russian]

\bibitem{Goryachev2007}
Goryachev, A.V., Mezhzil, E.K., Petrov, S.B., Syrov, V.A., Harlamov, A.V., Chivanov, S.V.: A way of ground testing objects of aircraft, subject to icing, and a device for its implementation. Patent RF, no. 2345345 (2007)

\bibitem{Klemenkov2008}
Klemenkov, G.P., Prihodko, Yu.M., Puzyrev, L.N., Haritonov, A.M.: Modelling of aircraft icing processes in aeroclimatic tubes. Thermophysics and Aeromechanics 15(4), 563--572  (2008) [in Russian]

\bibitem{Alekseenko2013}
Alekseenko, S.V., Prihodko, A.A.: Numerical simulation of icing cylinder and profile. Models review and calculation results. TsAGI Science Journal 44(6), 25--57 (2013) [in Russian]

\bibitem{Prihodko2014}
Prihodko, A.A., Alekseenko, S.V.: Numerical simulation of icing aerodynamic surfaces in the presence of large overcooled water drops. JETP Letters 40(19), 75--82 (2014) [in Russian]

\bibitem{Hannat2014}
Hannat, R., Morency, F.: Numerical Validation of Conjugate Heat Transfer Method for Anti-/De-Icing Piccolo System. J. Aircr. 51(1), 104--116 (2014); DOI: 10.2514/1.c032078

\bibitem{Villalpando2016}
Villalpando, F., Reggio, M., Ilinca, A.: Prediction of ice accretion and anti-icing heating power on wind turbine blades using standard commercial software. Energy 114, 1041--1052 (2016); DOI: 10.1016/j.energy.2016.08.047

\bibitem{Lynch2001}
Lynch, F.T., Khodadoust, A.: Effects of ice accretions on aircraft aerodynamics. Prog. Aeosp. Sci. 37(8), 669--767 (2001); DOI: 10.1016/s0376-0421(01)00018-5

\bibitem{Bragg2005}
Bragg, M.B., Broeren, A.P., Blumenthal, L.A.: Iced-airfoil aerodynamics. Prog. Aeosp. Sci. 41(5), 323--362 (2005); DOI: 10.4271/2003-01-2098

\bibitem{Modorskii_2_2016}
Modorskii, V.Ya., Sipatov, A.M., Babushkina, A.V., Kolodyazhny, D.Yu., Nagorny, V.S.: Modeling technique for the process of liquid film disintegration. In: Vasily Fomin (Ed.) ICMAR 2016, AIP Conference Proceedings, vol. 1770, art. no. 030109 (2016); DOI: 10.1063/1.4964051

\bibitem{Gaynutdinova2015}
Gaynutdinova, D.F., Modorsky, V.Y., Masich, G.F.: Infrastructure of Data Distributed Processing in High-Speed Process Research Based on Hydroelasticity Tasks. In: Sloot, P. (Ed.) YSC, Procedia Computer Science, vol. 66, pp. 556--563 (2015); DOI: 10.1016/j.procs.2015.11.063

\bibitem{Modorskii_3_2016}
Modorskii, V.Y., Gaynutdinova, D.F., Gergel, V.P., Barkalov, K.A.: Optimization in Design of Scientific Products for Purposes of Cavitation Problems. In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings, vol. 1738, art. no. 400013 (2016); DOI: 10.1063/1.4952201

\bibitem{Modorskii2002}
Modorskii, V.Y., Sokolkin, Y.V.: Dynamic Behavior of a thick-walled cylinder under pressurization. Izv. Vyss. Uchebnykh Zaved. Aviats. Tek. 4, 14--16 (2002)

\bibitem{Kozlova2010}
Kozlova, A.V., Modorskii, V.Y., Ponik, A.N.: Modeling of Cooling Processes in the Variable Section Channel of a Gas Conduit. Russian Aeronautics 53(4), 401--407 (2010); DOI: 10.3103/s1068799810040057

\bibitem{Strongin2000}
Strongin, R.G., Sergeyev, Y.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000); DOI: 10.1007/978-1-4615-4677-1

\bibitem{Barkalov2002}
Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)

\bibitem{Gergel2005}
Gergel, V.P., Strongin, R.G.: Parallel computing for globally optimal decision making on cluster systems. Future Gener. Comput. Syst. 21(5), 673--678 (2005); DOI: 10.1016/j.future.2004.05.007

\bibitem{Strongin2013}
Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013);  DOI: 10.1007/978-1-4614-8042-6

\bibitem{Barkalov2010}
Barkalov, K., Ryabov, V., Sidorov, S.: Parallel Scalable Algorithms with Mixed Local-Global Strategy for Global Optimization Problems. In: Hsu, Ch., Malyshkin, V. (Eds.) MTPP 2010, LNCS, vol. 6083, pp. 232--240. Springer, Heidelberg (2010); DOI: 10.1007/978-3-642-14822-4$\_$26

\bibitem{Barkalov2015}
Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi Coprocessor for Solving Global Optimization Problems. In: Malyshkin, V. (Ed.) PaCT 2015, LNCS, vol. 9251, pp. 307--318. Springer, Heidelberg (2015); DOI: 10.1007/978-3-319-21909-7$\_$31

\bibitem{Barkalov2016}
Barkalov, K., Gergel, V.: Parallel Global Optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016); DOI: 10.1007/s10898-016-0411-y

\bibitem{Gergel2016}
Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster. In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings, vol. 1738, art. no. 400006 (2016); DOI: 10.1063/1.4952194

\bibitem{Gergel2015}
Gergel, V., Sidorov, S.: A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems. In: Malyshkin, V. (Ed.) PaCT 2015, LNCS, vol. 9251, pp. 505--515. Springer, Heidelberg (2015); DOI: 10.1007/978-3-319-21909-7$\_$49

%\bibitem{jour} Smith, T.F., Waterman, M.S.: Identification of Common Molecular Subsequences. J. Mol. Biol. 147, 195--197 (1981)

%\bibitem{lncschap} May, P., Ehrlich, H.C., Steinke, T.: ZIB Structure Prediction Pipeline: Composing a Complex Biological Workflow through Web Services. In: Nagel, W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006. LNCS, vol. 4128, pp. 1148--1158. Springer, Heidelberg (2006)

%\bibitem{book} Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing Infrastructure. Morgan Kaufmann, San Francisco (1999)

%\bibitem{proceeding1} Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid Information Services for Distributed Resource Sharing. In: 10th IEEE International Symposium on High Performance Distributed Computing, pp. 181--184. IEEE Press, New York (2001)

%\bibitem{proceeding2} Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The Physiology of the Grid: an Open Grid Services Architecture for Distributed Systems Integration. Technical report, Global Grid Forum (2002)

%\bibitem{url} National Center for Biotechnology Information, \url{http://www.ncbi.nlm.nih.gov}

\end{thebibliography}

\end{document}

### 2017-09-14 2017/ParOptBook/Chapter4.tex
parOptBook: update chapter 4

@@
 %\usepackage[bottom]{footmisc}% places footnotes at page bottom
+%\usepackage{mathtools}
+%\mathtoolsset{showonlyrefs}
 %
@@
 be considered.
+Let $\varphi(x)$ and $g_j(x)\le 0,1\le j\le m$ be the real functions defined in a closed interval $[a,b]$ of the real axis.
+It is requires to find the values $x^*$ satisfying the condition
+\begin{equation}
+  \label{eq4:problem}
+  \varphi(x^*)=\min\{\varphi(x):x\in [a,b], g_j(x)\le 0,1\le j\le m\}
+\end{equation}
+
+As before, the objective function $\varphi$ (hereinafter denoted also as $g_{m+1}$) and the left parts of the constraints $g_j(x)\le 0,1\le j\le m$ re assumed to satisfy \emph{Lipschitz condition} with the corresponding constants $L_j,1\le j\le m+1$, namely:
+\begin{equation}
+  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2
+\end{equation}
+
+\example{
+\label{ex4:problem}
+Let us consider a problem of type \eqref{eq4:problem} where $x\in [0.6;2.2],m=2$,
+\begin{gather}
+  \varphi(x)=\cos(18x-3)\sin(10x-7)+1.5, \\
+  g_1(x)=\exp(-x/2)\sin(6x-1.5), \\
+  g_2(x)=|x|\sin(2\pi x-0.5).
+\end{gather}
+he exact solution of the problem is $x^*=2.0795,\varphi(x^*)=0.565$. The graphs of the functions $g_1(x),g_2(x)$,and $\varphi(x)$ are presented in Fig. \ref{fig:4_1}.
+\begin{figure}[h]
+  \label{fig:4_1}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_1.png}
+  \caption{One-dimensional global optimization problem}
+\end{figure}
+}
+
+\section{Partial computability and index scheme of the account of constraints}
+Problem \eqref{eq4:problem} can be considered in the statement when each function $g_j,1\le j\le m+1$ is defined and computable only in the corresponding subdomains $Q_j\subset [a,b]$ where
+\begin{equation}
+  \label{eq4:q}
+  Q_1=[a,b],Q_(j+1)=\{x\in Q_j:g_j(x)\le 0\},1\le j\le m.
+\end{equation}
+
+For example, in the optimal design problems some characteristics of the technical systems may appear to be undefined if the conditions of the system functioning represented by a part of the constraints of problem \eqref{eq4:problem} are not fulfilled.
+
+Taking into account conditions \eqref{eq4:q}, initial problem \eqref{eq4:problem} can be represented in the form
+\begin{equation}
+  \varphi(x^*)=\min\{g_{m+1}(x):x\in Q_{m+1}\}.
+\end{equation}
+
+For the purposes of further treatment, let us introduce a classification of the points $x$ in the search domain $[a,b]$ using an index $\nu=\nu(x)$ where $\nu-1$ is the number of constraints, which are fulfilled in this point.
+Formally, the index $\nu$ is defined by the conditions
+\begin{equation}
+  g_j(x)\le 0,1\le j \le \nu-1,g_\nu(x)>0,
+\end{equation}
+(the last inequality is inessential  if $\nu=m+1$) and it satisfies the inequalities
+\begin{displaymath}
+  1\le\nu=\nu(x)\le m+1.
+\end{displaymath}
+The problem of example \ref{ex4:problem} assuming a partial computability of the functions is presented in Fig.
+\ref{fig:4_2}.
+The arcs of the restrictions $g_1(x),\:g_2(x)$, and of the objective function $\varphi(x)$ defined in the corresponding subdomains $Q_j$ from \eqref{eq4:q} are shown.
+The point $x^1$ with the index $\nu(x^1)=1$, the point $x^2$ with the index $\nu(x^2)=2$, and the point $x^3$ with the index $\nu(x^3)=3$ are shown also.
+
+\begin{figure}[h]
+  \label{fig:4_2}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_2.png}
+  \caption{A problem with partially computable functions}
+\end{figure}
+
+This classification generates the function
+\begin{equation}
+  \label{eq4:fx}
+  f(x)=g_\nu(x),\nu=\nu(x),
+\end{equation}
+defined and computable everywhere in $[a,b]$.
+Its value at a point $x$ is either the value of the left part of the constraint violated at this point (the case $\nu\le m$)  or the value of the objective function (the case $\nu=m+1$).
+Therefore, the determination of the value of $f(x),\:x\in[a,b]$  is reduced to a successive computations of the values of $g_j(x), 1\le j\le \nu=\nu(x)$, i.e., the next value $g_{j+1}(x)$ is calculated in the case when $g_j(x)\le 0$ only.
+The computational process finishes either as a result of the fulfillment of the inequality $g_j(x)>0$ or as a result of achievement of the value $\nu(x)=m+1$.
+
+The described procedure called a \emph{trial} at the point $x$ forms the index $\nu$ of this point automatically. The pair of values
+\begin{equation}
+  z=f(x)=g_\nu(x),\nu=\nu(x)
+\end{equation}
+generated by the trial at the point $x\in[a,b]$ is called the \emph{trial result}.
+
+The graph of the function $f(x)$ from \eqref{eq4:fx}, which consists of the arcs of the constraints $g_1(x),\:g_2(x)$, and of the objective function $\varphi(x)$ taken from Example \ref{ex4:problem} is shown in Fig.
+\ref{fig:4_3}.
+
+\begin{figure}[h]
+  \label{fig:4_3}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_3.png}
+  \caption{An «index» function}
+\end{figure}
 

### 2017-09-17 2017/ParOptBook/Chapter4.tex
parOptBook: update chapter 4

@@
 \begin{equation}
+  \label{eq4:problem2}
   \varphi(x^*)=\min\{g_{m+1}(x):x\in Q_{m+1}\}.
@@
 \begin{equation}
+  \label{eq4:condition}
   g_j(x)\le 0,1\le j \le \nu-1,g_\nu(x)>0,
@@
 \begin{equation}
+  \label{eq4:trial}
   z=f(x)=g_\nu(x),\nu=\nu(x)
@@
 
+Since in the general case the problem \eqref{eq4:problem2} may have no solution (i.e., the feasible subdomain $Q_{m+1}$ might appear to be empty because of the incompatibility of the constrains) an auxiliary problem always having a solution is associated with it.
+Since conditions \eqref{eq4:condition} are equivalent to the conditions
+\begin{displaymath}
+  x\in Q_\nu,x\not\in Q_{\nu+1},
+\end{displaymath}
+this auxiliary problem can be written in the form
+\begin{equation}
+  \label{eq4:problem3}
+  g_M^*=g_M(x_M^*)=\min\{g_M(x):x\in Q_M\},
+\end{equation}
+where $M$ is the greatest possible value of the index i. e.
+\begin{equation}
+  1\le M=\max\{\nu(x):x\in[a,b]\}\le m+1
+\end{equation}
+Since the set $Q_M$ is always nonempty, problem \eqref{eq4:problem3} always has a solution.
+When $M=m+1$ the solution $x^*=x^*_m+1$  is also the solution of the initial problem \eqref{eq4:problem}.
+When $M<m+1$ the inequality $g_M^*<0$ fulfilled necessarily can be used as an indicator of the incompatibility of the constraints.
+
+The main idea of the index approach is to reduce the constrained problem \eqref{eq4:problem3} to a unconstrained one
+\begin{displaymath}
+  \psi(x^*)=\min\{\psi(x):x\in[a,b]\},
+\end{displaymath}
+where
+\begin{equation}
+  \psi(x)=
+  \begin{cases}
+    g_\nu(x)/L_\nu, & \nu < M, \\
+    (g_M-g_M^*), & \nu=M.
+  \end{cases}
+\end{equation}
+
+As a result, the arcs of $\psi(x)$ will be Lipschitzian with the constant $L=1$ in each subdomain $Q_\nu,1\le\nu\le M$.
+A graph of the function $\psi(x)$ for the example (\ref{ex4:problem}) is presented in Fig.
+\ref{fig:4_4}.
+This new function will have discontinuities of the first type at the boundary points of the subdomains $Q_\nu$ from \eqref{eq4:q}.
+Nevertheless, one can estimate the global minimizer $x^*_M$ using the results of $k$ trials for \eqref{eq4:trial} at the points $x_1,\dots, x_k$ in $[a,b]$ (see \cite{3}).
+
+Indeed, as follows from Lipschitz condition
+\begin{equation}
+  \label{eq4:lip_opt_est}
+  x^*_M\in\{x\in[a,b]:|x-x^i|\ge\psi(x^i),1\le i\le k\}.
+\end{equation}
+
+For example, a case for $k=4$ is presented in Fig.
+\ref{fig:4_4}.
+The union of the segments highlighted with a bold line in Fig.
+\ref{fig:4_4} is a complement of the set from the right hand side of \eqref{eq4:lip_opt_est} and does not include the optimum point.
+\begin{figure}[h]
+  \label{fig:4_4}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_4.png}
+  \caption{An estimate of the optimum point}
+\end{figure}
+
+Note that the greatest index $M$, the values of Lipschitz constants $L_\nu, 1\le \nu\le M$, and the value of $g_M^*$ are the unknowns.
+However, these problems can be overcome by means of using instead of them the adaptive estimates of these values obtained in the course of solving the problem on the base of trial results.
+
+Having introduced all necessary concepts, let us turn to the description of the index algorithm.
+
+\section{The index global search algorithm}
+The first trial is executed at an arbitrary internal point $x_1\in(a,b)$. The selection of any subsequent trial point $x^{k+1},k\ge 1$ is determined by the following rules.
+
+\emph{Rule 1.} Renumber the points $x^1,...,x^k$ of the preceding trials by the lower indices in ascending order of coordinate values, i.e.
+\begin{equation}
+  \label{eq4:search_data}
+  0=x_0<x_1<\dots <x_k<x_{k+1}=1,
+\end{equation}
+and juxtapose to them the values $z_i=g_\nu(y(x_i)), \; \nu=\nu(x_i), \; 1
+\leq i \leq k$, from \eqref{eq4:fx}; the points $x_0=0$ and
+$x_{k+1}=1$ are introduced additionally for the convenience  of further notation (the values $z_0$ and $z_{k+1}$ are not defined).
+
+\emph{Rule 2.} Classify the indices $i, \; 1 \leq i \leq k$, of the trial points from set \eqref{eq4:search_data} according to the number of the problem constraints fulfilled at these points, by constructing the sets
+\begin{equation}
+  I_\nu =\left\{i:1 \leq i \leq k, \; \nu=\nu(x_i) \right\}, \; 1 \leq \nu \leq m+1,
+\end{equation}
+
+containing the numbers of all the points $x_i, \; 1 \leq i \leq k$, with
+the same values of $\nu$. The end points $x_0=0$ and $x_{k+1}=1$ are
+interpreted as the ones having indices equal to zero. An additional set,
+$I_0=\left\{0,k+1\right\}$, corresponds to them.
+
+\emph{Rule 3.} Determine the maximum value of the index:
+\begin{equation}
+  M=\max\left\{\nu(x_i), \; 1 \leq i \leq k \right \}.
+\end{equation}
+
+\emph{Rule 4.} Compute the current lower estimates,
+\begin{equation}
+  \label{eq4:lip_est}
+  \mu = \max\left\{ \frac{\left|z_i-z_j\right|}{ x_i - x_j }, \; i,j \in I_\nu, \; i>j \right\},
+\end{equation}
+for the unknown Lipschitz constants $L_\nu$ of the functions $g_\nu(y),1
+\leq \nu \leq m+1$. If a set $I_\nu$ contains less than two elements, or
+if $\mu_\nu$ is equal to zero, then assume $\mu_\nu=1$.
+As follows from \eqref{eq4:lip_est}, the estimates $\mu_\nu$ are non-decreasing starting from the moment when \eqref{eq4:lip_est} generates a positive value $\mu_\nu$.
+
+\emph{Rule 5.} For all nonempty sets $I_\nu, \; 1 \leq \nu \leq m+1$, compute the estimates
+\begin{equation}
+  z_\nu^\ast = \left\{
+  \begin{array}{lr}
+    0, & \nu < M,\\
+    \min\{ g_\nu(y(x_i): i\in I_\nu \}, & \nu = M.
+  \end{array}
+  \right.
+\end{equation}
+
+\emph{Rule 6.} For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} $R(i)$:
+\[
+R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
+\]
+\begin{equation}%\label{eq:14}
+R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
+\end{equation}
+\[
+R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
+\]
+where $\Delta_i=(x_i - x_{i-1})^{1/N}$. The values $r_\nu > 1, \; 1 \leq
+\nu \leq m+1,$ are parameters of the algorithm. An appropriate selection
+of $r_\nu$ allows to consider the product $r_\nu \mu_\nu$ as an estimate
+of the H{\"o}lder constants $H_\nu, \; 1 \leq \nu \leq m+1$.
+
+\emph{Rule 7.} Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
+\begin{equation}\label{MaxR}
+R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
+\end{equation}
+
+Rule 7. Make the next trial at the midpoint of the interval
+$(x_{t-1},x_t)$ if the indices of the points $x_{t-1}$ and $x_t$  are not
+the same, i.e.
+\[
+x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
+\]
+Otherwise, make the trial at the point
+\begin{equation}%\label{eq:142}
+x^{k+1} = \frac{x_t+x_{t-1}}{2} - \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
+\end{equation}
+
+We may take as termination condition the inequality $\Delta_t \leq
+\epsilon$, where $t$ is from (\ref{MaxR}) and $\epsilon>0$ is the
+predefined accuracy.
+
+Various modifications of this algorithm and the corresponding theory of
+convergence are given in
+\cite{Strongin2000,Barkalov2002,Gergel2005,Strongin2013}.
+
+The algorithm considered above is very flexible and allows for an
+efficient parallelization for shared memory, for distributed memory, and
+for accelerators \cite{Barkalov2010,Barkalov2015,Barkalov2016,Gergel2016}.
+
+\section{Index algorithm taking into account the existence of -reserved solutions}
+
+\section*{Index method with adaptive order of checking for constraints}
+According to the rules of the index method considered in the preceding subsections, every trial includes the successive check of the fulfillment of the problem constraints in the corresponding point of the search domain.
+The finding of the first violated constraint terminates the trial and initiates the transition to the next iteration.
+If a numeration of the constraints in the increasing order by the computational costs of their checking is introduced (i.e., if the simpler constraints are checked first) one can reduce essentially the overall computational costs in the course of optimization.
+The above is important for the problems, in which the estimating of the function value requires a considerable computational resources connected with a necessity of the numerical modelling of the behavior of the systems described by these functions (as it takes place, for example, in numerous problems of the optimal design).
+
+Obviously, the computational expenditures of the trial execution depend on the order of the problem functions’ computation essentially.
+If the calculation time of any problem is independent on the iteration execution point, the less the number of the problem functions (constraints and objective function) calculated at a point x, the less the trial execution costs in this point.
+
+For many applied problems, it is naturally to assume that a constraint violated at a point in the search domain, will be violated in some neighborhood of this point as well.
+In this case, it might be useful to begin the next trials in the points of this neighborhood with checking the fulfillment of this particular constraint.
+In the scope of the above, it seems reasonable to design a global optimization algorithm allowing the modification of the constraint  check order taking into account the information obtained during the search.
+The check order design will be oriented onto the constraint violated at an iteration point to be found as early as possible.
+
+\section{Parallel computations for the multiextremal optimization problems with nonlinear constraints}
+
+\section{A brief conclusion}
+Constrained global optimization problems are under consideration.
+The method of penalty functions is one of the most popular numerical methods for solving the problems of this kind.
+Nevertheless, the method of penalty functions has a number of disadvantages.
+A new method for reduction of a constrained problem to an unconstrained one, so called \emph{index method}, has been considered in this chapter.
+This method, unlike the classical penalty function one, allows avoiding the problem of the parameter selection (like the penalty coefficients for the constraints).
+It is also allows solving the problems with partially defined objective function and constraints.
+This is quite natural for many applied problems, especially, in the optimal design of technical systems, because if some conditions of operation are not met, then some other characteristics of the system performance may not be defined.
+
+The issues of convergence acceleration of the index method have been considered also.
+Two approaches for the convergence acceleration have been proposed.
+The first of them is based on concept of $\varepsilon$-reserved solution to the constrained problem.
+The role of $\varepsilon$-reserved solution in constrained global optimization problems is somewhat similar to the regularity requirements in classical nonlinear programming problems.
+The second approach is based on adaptive order of checking for constraints.
+In the index method, every iteration performed at the corresponding point of the search domain includes checking for constraints of the problem at this point.
+When a violation is discovered for the first time, the current iteration is stopped, and the next iteration is initiated.
+However, in Lipschitz optimization problems a constraint violated at a point of the search domain is also violated in a certain neighborhood of this point.
+In this case, to reduce the number of checks, it is expedient to begin iterations in the neighborhood of this point with checking for the violated constraint.
+Thus, the iteration can be completed at a lower cost.
+
+For all proposed algorithms, the results of numerical experiments are presented. The results confirm the theoretical statements about convergence acceleration.
+
 %\end{document}

### 2017-09-18 2017/ParOptBook/Chapter4.tex
ParOptBook: update Chapter4.tex

@@
 
-%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\documentclass[graybox]{svmult}
-%
-%% choose options for [] as required from the list
-%% in the Reference Guide
-%
-%\usepackage{mathptmx}       % selects Times Roman as basic font
-%\usepackage{helvet}         % selects Helvetica as sans-serif font
-%\usepackage{courier}        % selects Courier as typewriter font
-%\usepackage{type1cm}        % activate if the above 3 fonts are
-%                           % not available on your system
-%
-%\usepackage{makeidx}         % allows index generation
-%\usepackage{graphicx}        % standard LaTeX graphics tool
-%                             % when including figure files
-%\usepackage{multicol}        % used for the two-column index
-%\usepackage[bottom]{footmisc}% places footnotes at page bottom
-%\usepackage{mathtools}
-%\mathtoolsset{showonlyrefs}
-%
-%% see the list of further useful packages
-%% in the Reference Guide
-%
-%\makeindex             % used for the subject index
-%                       % please use the style svind.ist with
-%                       % your makeindex program
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\begin{document}
+% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[graybox]{svmult}
+
+% choose options for [] as required from the list
+% in the Reference Guide
+
+\usepackage{mathptmx}       % selects Times Roman as basic font
+\usepackage{helvet}         % selects Helvetica as sans-serif font
+\usepackage{courier}        % selects Courier as typewriter font
+\usepackage{type1cm}        % activate if the above 3 fonts are
+                           % not available on your system
+
+\usepackage{makeidx}         % allows index generation
+\usepackage{graphicx}        % standard LaTeX graphics tool
+                             % when including figure files
+\usepackage{multicol}        % used for the two-column index
+\usepackage[bottom]{footmisc}% places footnotes at page bottom
+\usepackage{mathtools}
+\mathtoolsset{showonlyrefs}
+
+% see the list of further useful packages
+% in the Reference Guide
+
+\makeindex             % used for the subject index
+                       % please use the style svind.ist with
+                       % your makeindex program
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{document}
 
@@
 \begin{equation}
-  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2
+  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2\in [a,b]
 \end{equation}
@@
 \]
-\begin{equation}%\label{eq:14}
+\begin{equation}
+\label{eq4:characteristic}
 R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
@@
 \]
-where $\Delta_i=(x_i - x_{i-1})^{1/N}$. The values $r_\nu > 1, \; 1 \leq
-\nu \leq m+1,$ are parameters of the algorithm. An appropriate selection
-of $r_\nu$ allows to consider the product $r_\nu \mu_\nu$ as an estimate
-of the H{\"o}lder constants $H_\nu, \; 1 \leq \nu \leq m+1$.
-
+\[
+\Delta_i=x_i - x_{i-1}
+\]
+The values $r_\nu>1, 1\le\nu\le m+1$ are the parameters of the algorithm.
+An appropriate choice of the values $r_\nu$ allows using the products $r_\nu\mu_\nu$ as the estimates of Lipschitz constants $L_\nu, 1\le\nu\le m+1$.
 \emph{Rule 7.} Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
-\begin{equation}\label{MaxR}
+\begin{equation}
+\label{eq4:MaxR}
 R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
 \end{equation}
-
-Rule 7. Make the next trial at the midpoint of the interval
-$(x_{t-1},x_t)$ if the indices of the points $x_{t-1}$ and $x_t$  are not
-the same, i.e.
+\emph{Rule 8.} Make the next trial at the midpoint of the interval
+$(x_{t-1},x_t)$ if the indices of the points $x_{t-1}$ and $x_t$  are not the same, i.e.
 \[
@@
 Otherwise, make the trial at the point
-\begin{equation}%\label{eq:142}
-x^{k+1} = \frac{x_t+x_{t-1}}{2} - \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
+\begin{equation}
+\label{eq4:next_point}
+x^{k+1} = \frac{x_t+x_{t-1}}{2} - \frac{z_t-z_{t-1}}{2r_\nu\mu},\; \nu=\nu(x_{t-1})=\nu(x_t).
 \end{equation}
 
-We may take as termination condition the inequality $\Delta_t \leq
-\epsilon$, where $t$ is from (\ref{MaxR}) and $\epsilon>0$ is the
-predefined accuracy.
+The described rules can be supplemented with the termination condition, finalizing the iterations of the method if
+\begin{equation}
+\label{eq4:stop_cond}
+  x_t - x_{t-1}\le\epsilon
+\end{equation}
+where t from \eqref{eq4:MaxR} and $\epsilon>0$ is the predefined accuracy of the optimization.
 
-Various modifications of this algorithm and the corresponding theory of
-convergence are given in
-\cite{Strongin2000,Barkalov2002,Gergel2005,Strongin2013}.
+Let us formulate the conditions of convergence for the algorithm in the form of the following theorem.
+\begin{theorem}
+Let the considered index algorithm be applied to the solving of problem \eqref{eq4:problem} and the following conditions be satisfied:
+\begin{enumerate}
+  \item each function $gj, 1\le j\le m+1$ satisfies Lipschitz condition with the constant $L_j$ in the interval $[a,b]$, i.e.,
+  \[
+  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2 \in[a,b]
+  \]
+  \item the following inequalities hold for the quantities $\mu_\nu$ from \eqref{eq4:lip_est} starting from some step
+  \[
+  \r_\nu\mu_\nu > 2L_\nu,1\le\nu\\le m+1.
+  \]
+\end{enumerate}
 
-The algorithm considered above is very flexible and allows for an
-efficient parallelization for shared memory, for distributed memory, and
-for accelerators \cite{Barkalov2010,Barkalov2015,Barkalov2016,Gergel2016}.
+Then the set of the limit points of the sequence $\{x_k\}$ generated by the index algorithm coincides with the set of solutions of problem \eqref{eq4:problem} with $\epsilon=0$ and termination criterion \eqref{eq4:stop_cond}.
+In addition, the index of each limit point equals to $M$ and the convergence to any limit point $\overline{x}$  is bilateral if $\overline{x}\not=a$ and $\overline{x}\not=b$.
+\end{theorem}
+
+This theorem is given without a proof, which can be found, for example, in \cite{}[1, 2].
+Various modifications of this algorithm and the corresponding theory of convergence are given in \cite{}[5–10].
 
@@
 
-%\end{document}
+\end{document}

### 2017-09-18 2017/ParOptBook/Chapter4.tex
parOptBook: update chapter 4

@@
   \includegraphics[width=0.8\textwidth]{figures/4_3.png}
-  \caption{An «index» function}
+  \caption{An ''index'' function}
 \end{figure}
@@
   \[
-  \r_\nu\mu_\nu > 2L_\nu,1\le\nu\\le m+1.
+  r_\nu\mu_\nu > 2L_\nu,1\le\nu\\le m+1.
   \]
@@
 
-\section{Index algorithm taking into account the existence of -reserved solutions}
+\textbf{The results of the experiment.} Let us consider a problem from Example (\ref{ex4:problem}) as an illustration.
+Under assumption of partial computability the arcs of the problem functions corresponding to the domains $Q_j, 1\le j\le 3$, from \eqref{eq4:q} are presented in Fig.
+\ref{fig:4_6}.
 
-\section*{Index method with adaptive order of checking for constraints}
+Three methods have been used for solving this example:
+\begin{enumerate}
+
+  \item Scanning a uniform grid with the precision (distance between adjacent nodes) $\epsilon=10^{-5}$.
+  \item Penalty function method, i.e., the minimization of the function
+  \[
+  p(x)=\varphi(x)+(\max\{g_1(x),0\})^2+(\max\{g_2(x),0\})^2
+  \]
+  using Algorithm of Global Search.
+The coordinates of the trial points executed by AGS in the process of problem solving are marked by the vertical strokes in Fig.
+\ref{fig:4_5}.
+The coordinates of the trials executed in the close points are marked by a dark rectangle.
+  \item Index algorithm with the parameters $r_\nu=2, 1\le \nu \le 3$, and $\epsilon =10^{- 5}$.
+The trial point coordinates executed by the algorithm in the course of problem solving are marked by three rows of the vertical strokes in Fig.
+\ref{fig:4_6}.
+The strokes in the upper row correspond to the points with the index $\nu=1$, the ones of the second row --- to the points, the indices of which are equal to 2; and the points marked by the strokes in the bottom row are the feasible ones.
+The trial coordinates performed in the close points are marked by a dark rectangle.
+\end{enumerate}
+
+\begin{figure}[h]
+  \label{fig:4_5}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_5.png}
+  \caption{Minimization by means of Penalty Function Method}
+\end{figure}
+
+It is worth noting, that in the applied problems the estimates of the constraints as well as of the objective functions require considerable computation resources very often.
+The results presented in Table \ref{tab4:exp_results} confirm the effectiveness of the index scheme of the account of
+constraints.
+
+\begin{table}
+
+\begin{center}
+  \label{tab4:exp_results}
+  \begin{tabular}{| l | c | c | c| }
+    \hline
+    & $k_1$ & $k_2$ & $k_3$ \\ \hline
+    Scanning a uniform grid & 160000 & 90280 & 56476 \\ \hline
+    Penalty Function Method &  375 & 375 & 375 \\ \hline
+    Index method & 63 & 49 & 35 \\ \hline
+  \end{tabular}
+\end{center}
+\end{table}
+
+Here $k_i$ is the number of computations of the $i$ problem function, i.e., of the function $g_i(x )$.
+\begin{figure}[h]
+  \label{fig:4_6}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_6.png}
+  \caption{The minimization of an ''index'' function}
+\end{figure}
+
+\section{Index algorithm taking into account the existence of $\epsilon$-reserved solutions}
+Let us return to the consideration of a one-dimensional problem of the kind \eqref{eq4:problem}
+\begin{displaymath}
+  \varphi(x^*)=\min\{\varphi(x):x\in [a,b], g_j(x)\le 0,1\le j\le m\}
+\end{displaymath}
+In the case when the objective function and the left-side parts of the constraints are partially computable, it is necessary to use the form \eqref{eq4:problem2}.
+\example
+{
+Let us consider problem \eqref{eq4:problem} with $x\in[0,6,2,2]$, $m=3$ as an illustration:
+}
+
+\begin{figure}[h]
+  \label{fig:4_7}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_7.jpg}
+  \caption{The solution of a problem by Index Method}
+\end{figure}
+
+\begin{figure}[h]
+  \label{fig:4_8}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_8.jpg}
+  \caption{The problem solving by Index Method with the $\epsilon$-reservs}
+\end{figure}
+
+\section{Index method with adaptive order of checking for constraints}
 According to the rules of the index method considered in the preceding subsections, every trial includes the successive check of the fulfillment of the problem constraints in the corresponding point of the search domain.
@@
 
-\section{A brief conclusion}
+\section*{A brief conclusion}
 Constrained global optimization problems are under consideration.

### 2017-09-19 2017/ParOptBook/Chapter4.tex
parOptBook: fix some infrastructure things in Chapter4

@@
 \usepackage[bottom]{footmisc}% places footnotes at page bottom
+
+%custom package
 \usepackage{mathtools}
-\mathtoolsset{showonlyrefs}
 
@@
 Let us consider a problem of type \eqref{eq4:problem} where $x\in [0.6;2.2],m=2$,
-\begin{gather}
+\begin{gather*}
   \varphi(x)=\cos(18x-3)\sin(10x-7)+1.5, \\
@@
   g_2(x)=|x|\sin(2\pi x-0.5).
-\end{gather}
+\end{gather*}
 he exact solution of the problem is $x^*=2.0795,\varphi(x^*)=0.565$. The graphs of the functions $g_1(x),g_2(x)$,and $\varphi(x)$ are presented in Fig. \ref{fig:4_1}.
@@
 \emph{Rule 6.} For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} $R(i)$:
-\[
-R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
-\]
-\begin{equation}
+\begin{gather}
   \label{eq4:characteristic}
-R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
-\end{equation}
-\[
-R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
-\]
-\[
-\Delta_i=x_i - x_{i-1}
-\]
+  R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}), \nonumber \\
+  R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),\\
+  R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i), \nonumber \\
+  \Delta_i=x_i - x_{i-1} \nonumber
+\end{gather}
 The values $r_\nu>1, 1\le\nu\le m+1$ are the parameters of the algorithm.
 An appropriate choice of the values $r_\nu$ allows using the products $r_\nu\mu_\nu$ as the estimates of Lipschitz constants $L_\nu, 1\le\nu\le m+1$.
+
 \emph{Rule 7.} Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
@@
 \end{equation}
+
 \emph{Rule 8.} Make the next trial at the midpoint of the interval
@@
 \begin{table}
-
+  \caption{}
 \begin{center}

### 2017-09-20 2017/ParOptBook/Chapter4.tex
parOptBook: update chapter 4

@@
 \begin{equation}
+  \label{eq4:mu_const}
   z_\nu^\ast = \left\{
@@
 Let us consider problem \eqref{eq4:problem} with $x\in[0,6,2,2]$, $m=3$ as an illustration:
+\begin{gather}
+\varphi(x)=\cos(18x-3)\sin(10x-7)+1, \\
+g_1(x)=\exp(-x/2)\sin(6x-1.5) \\
+g_2(x)=\sin(4x-2.2)+\cos(6x-2.9) \\
+g_3(x)=|x|\sin(2\pi x - 0.5).
+\end{gather}
 }
-
 \begin{figure}[h]
@@
 \end{figure}
+The functions $g_j , 1\le j\le 4$, are pictured in Fig. \ref{fig:4_7}. Assuming a partial calculability, the arcs of
+the functions corresponding to the domains $Q_j ,1\le j\le 4$, from \eqref{eq4:q} are presented in Fig. \ref{fig:4_8}.
 
@@
 \end{figure}
+
+Index algorithm described in Sec.
+3.3 was used to solve this example with $x_1=(a+b)/2$, $r=2$, and $\varepsilon=10^{-5}$.
+The coordinates of 102 trial points executed by the algorithm in the process of problem solving are marked by four rows of the vertical strokes in Fig.
+\ref{fig:4_7}.
+The strokes of the upper row correspond to the points with the index $\nu=1$, the ones of the second and the third rows --- to the points, the indices of which are equal to 2 and 3, respectively; and the points
+marked by the strokes of the bottom row are feasible.
+
+The coordinates of the trials executed in the close points are marked by a dark rectangle.
+The values of the functions $g_1 , g_2 , g_3$ , and $\varphi=g_4$ have been computed $k_1=102$, $k_2=80, k_3=64,$ and $k_4=26$ times, respectively.
+As one can see in Fig.
+\ref{fig:4_7}, the trial points concentration takes place not only in the vicinity of the point of the global optimum $x^*=2,07957$, but also in the vicinities of the inadmissible boundary points of the domains $Q_j$ from \eqref{eq4:q}.
+This effect can be reduced by the substitution of estimates \eqref{eq4:mu_const} by the estimates
+\begin{equation}
+  \label{eq4:mu_const}
+  z_\nu^\ast = \left\{
+  \begin{array}{lr}
+    -\epsilon_\nu, & \nu < M,\\
+    \min\{ g_\nu(y(x_i): i\in I_\nu \}, & \nu = M.
+  \end{array}
+  \right.
+\end{equation}
+where $\epsilon_R=(\epsilon_1 ,\dots, \epsilon_m)$ is a predefined vector with positive components.
+Thus, $z^*=0$ if there exist the points $x_i,1\le i\le k$ from series \eqref{eq4:search_data} with the index greater than $\nu$.
+As a consequence of this substitution the values of the characteristics $R(i)$ from \eqref{eq4:characteristic} will become lower (because of the subtraction of $\epsilon_\nu$) if
+\begin{equation}
+  \nu=\max\{\nu(x_{i-1},\nu(x_i)\}<\max\{\nu(x_j):1\le j\le k\}\le V
+\end{equation}
+that makes executing the trials in the corresponding intervals  ($x_{i-1}$ , $x_i$ ) less probable. In all other aspects, the algorithm remains the same.
+
+This \emph{prima facie} mechanical trick has a well-established theoretical basis considered below.
+
+\begin{definition}
+  The point $x_\epsilon$ is called an $\epsilon$-reserved solution of problem \eqref{eq4:problem} if the condition
+  \begin{equation}
+    \varphi(x_\epsilon)=\min\{\varphi(x):x\in [a,b], g_j(x)\le -\epsilon_j,1\le j\le m\}
+  \end{equation}
+  meets where $\epsilon_1, \dots,\epsilon_m$ are the positive values of reserve for each constraint. Let us introduce into consideration also the set
+  \begin{equation}
+  X_\epsilon = \{x\in[a,b]:g_j(x)\le 0, 1\le j\le m, \varphi(x)\le\varphi(x_\epsilon)\}
+  \end{equation}
+  of all feasible solutions of the problem, which are no worse with respect to the objective function value than the $\epsilon$-reserved solution.
+\end{definition}
+
+The existence of the $\epsilon$-reserved solution of problem \eqref{eq4:problem} can be interpreted as some analogue of the conditions of regularity in the classical nonlinear programming problems.
+The applied role of this condition should be noted also.
+Even if the exact solution $x^*$ is known, its practical implementation is possible as some approximation to $x^*$ only.
+Therefore, the existence of the feasible points from $Q$ close to $x^*$ (with respect to the coordinates and to the objective
+function values) is important.
+The existence of the $\epsilon$-reserved solution guarantees the availability of such points.
+Moreover, the domain $X_\epsilon$ may play a role of the set of suitable approximations.
+
+The index algorithm with the $\epsilon$-reserves has been applied to the considered example with $\epsilon_i=0.2, 1\le i\le 3$.
+The search trial coordinates are marked by the vertical strokes in Fig.
+\ref{fig:4_8}.
+The solution has required 52 trials only and provided the same precision of the problem solution as in
+the case when the vector $\epsilon_R$ equals to zero.
+Moreover, the values of the functions $g_1$ , $g_2$ , $g 3$ , and $\varphi=g_4$ have been calculated $k_1 =52$, $k_2 =39$, $k_3 =38$, and $k_4 =25$ times, respectively.
+
+The conditions of convergence for the algorithm are presented in the following theorem (given without proof, see \cite{}).
+
+\begin{theorem}
+  Assume
+\end{theorem}
 

### 2017-09-23 2017/ParOptBook/Chapter4.tex
parOptBook: update chapter 4

@@
 
-% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[graybox]{svmult}
-
-% choose options for [] as required from the list
-% in the Reference Guide
-
-\usepackage{mathptmx}       % selects Times Roman as basic font
-\usepackage{helvet}         % selects Helvetica as sans-serif font
-\usepackage{courier}        % selects Courier as typewriter font
-\usepackage{type1cm}        % activate if the above 3 fonts are
-                           % not available on your system
-
-\usepackage{makeidx}         % allows index generation
-\usepackage{graphicx}        % standard LaTeX graphics tool
-                             % when including figure files
-\usepackage{multicol}        % used for the two-column index
-\usepackage[bottom]{footmisc}% places footnotes at page bottom
-
-%custom package
-\usepackage{mathtools}
-
-% see the list of further useful packages
-% in the Reference Guide
-
-\makeindex             % used for the subject index
-                       % please use the style svind.ist with
-                       % your makeindex program
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
+%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\documentclass[graybox]{svmult}
+%
+%% choose options for [] as required from the list
+%% in the Reference Guide
+%
+%\usepackage{mathptmx}       % selects Times Roman as basic font
+%\usepackage{helvet}         % selects Helvetica as sans-serif font
+%\usepackage{courier}        % selects Courier as typewriter font
+%\usepackage{type1cm}        % activate if the above 3 fonts are
+%                           % not available on your system
+%
+%\usepackage{makeidx}         % allows index generation
+%\usepackage{graphicx}        % standard LaTeX graphics tool
+%                             % when including figure files
+%\usepackage{multicol}        % used for the two-column index
+%\usepackage[bottom]{footmisc}% places footnotes at page bottom
+%
+%%custom package
+%\usepackage{mathtools}
+%
+%% see the list of further useful packages
+%% in the Reference Guide
+%
+%\makeindex             % used for the subject index
+%                       % please use the style svind.ist with
+%                       % your makeindex program
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%\begin{document}
 
@@
 Let us consider problem \eqref{eq4:problem} with $x\in[0,6,2,2]$, $m=3$ as an illustration:
-\begin{gather}
+\begin{gather*}
 \varphi(x)=\cos(18x-3)\sin(10x-7)+1, \\
@@
 g_3(x)=|x|\sin(2\pi x - 0.5).
-\end{gather}
+\end{gather*}
 }
@@
   \begin{equation}
+    \label{eq4:eps_res_problem}
     \varphi(x_\epsilon)=\min\{\varphi(x):x\in [a,b], g_j(x)\le -\epsilon_j,1\le j\le m\}
@@
   \begin{equation}
+    \label{eq4:x_epsilon}
     X_\epsilon = \{x\in[a,b]:g_j(x)\le 0, 1\le j\le m, \varphi(x)\le\varphi(x_\epsilon)\}
@@
 \begin{theorem}
-  Assume
+  Assume problem \eqref{eq4:problem} to have a solution $x^*$ and the following conditions to be satisfied:
+  \begin{enumerate}
+    \item each domain $Q_j ,1\le j\le m+1$ from \eqref{eq4:q} is a union of a finite number of positive length segments;
+    \item each function $g_j ,1\le j\le m+1$ allows Lipschitz extension $G_j (x)$ with the constant $L_j$ over the whole interval $[a,b]$, i.e.,
+    \begin{equation}
+      g_j(x)=G_j(x),x\in Q_j,1\le j\le m+1;
+    \end{equation}
+    \item the components $\epsilon_\nu$ of the reserve vector $\epsilon_R$ corresponding to the constraints being active at the absolute minimum point $x^*$ (i.e., the constraints, for which $g_\nu(x^*)=0$) satisfy the inequalities
+    \begin{equation}
+      \label{eq4:reserves}
+      0<2\epsilon_\nu<L_\nu(\beta-\alpha)
+    \end{equation}
+    where $\beta-\alpha$ is the length of the interval $[\alpha,\beta]\subset Q_{m + 1}$ containing the point $x^*$;
+    \item the components $\epsilon_\nu$ of the reserve vector $\epsilon_R$ corresponding to the constraints which are not active at the absolute minimum point $x^*$ (i.e., $g_\nu(x^* )<0)$ satisfy either inequalities \eqref{eq4:reserves} or the inequalities
+    \begin{equation}
+      0< \epsilon_\nu <|g_\nu (x^* )|;
+    \end{equation}
+  \end{enumerate}
+
+  Then:
+  \begin{enumerate}
+    \item the point $x^*$ is the limit point of the trial sequence $\{x_k \}$ generated by index algorithm for problem \eqref{eq4:problem} at $\epsilon =0$ in the termination condition \eqref{eq4:stop_cond};
+    \item any limit point $\overline x$ of the sequence ${x_k }$ is a solution of the problem \eqref{eq4:problem};
+    \item the convergence to the limit point $x$ is a bilateral one if $x\not=a$ and $x\not=b$.
+    If the conditions \eqref{eq4:reserves} and (4.27) are not satisfied for given vector of reserves $\epsilon_R$ but the
+    problem $\eqref{eq4:problem}$ has an $\epsilon$-reserved solution $x_\epsilon$ defined by \eqref{eq4:eps_res_problem}, then:
+
+    \item for any limit point $x$ of the sequence ${x_k}$ it is true that
+    \begin{displaymath}
+      \varphi(\overline x)=\inf\{\varphi(x^k):g_j(x^k)\le 0,\: 1\le j\le m,\: k=1,2,\dots\}\le\varphi(x_\epsilon);
+    \end{displaymath}
+    \item any limit point $\overline x$ is a boundary point of the set $X_\epsilon$ from \eqref{eq4:x_epsilon} if the function $\varphi(x)$ has no more than one point of local maximum in each of the isolated intervals making up the set $X_\epsilon$;
+    \item the above statement on the bilateral character of the convergence remains in force.
+  \end{enumerate}
 \end{theorem}
+
+As it has been demonstrated by the experiment with solving the example described above, the increase of the reserve values may result in an essential decrease of the trial concentration in the boundary points of non-feasible domains.
+The theoretical properties of the trial sequences generated by the index algorithm are determined by the following statement (see the proof in \cite{}[4] also).
+\begin{theorem}
+  Assume the convergence conditions for the index algorithm to be satisfied and the inequalities
+  \begin{displaymath}
+    g_\nu(x)\ge\delta_\nu,x\in[\alpha,\beta],\delta_\nu\ge 0
+  \end{displaymath}
+  to be satisfied in a subinterval $[\alpha,\beta]\subset Q_\nu \: 1\le\nu\le m$, the length of which meets the condition
+  \begin{displaymath}
+    \beta-\alpha>\frac{2(r_\nu-1)(\delta_\nu+\epsilon_\nu)}{r^2_\nu L_\nu}.
+  \end{displaymath}
+
+  Then for the density $P_{\alpha\beta}$ of the points from the sequence $\{x_k\}$ generated by the index algorithm while solving the problem \eqref{eq4:problem} there is the estimate
+  \begin{displaymath}
+    P_{\alpha\beta}<\frac{r^2_\nu L_\nu}{(r_\nu-1)(\delta_\nu+\epsilon_\nu)}.
+  \end{displaymath}
+\end{theorem}
+
+\textbf{Selection of components of the vector of reserves.} A problem of the choice of concrete component values in the vector of reserves $\epsilon_R$ arises when applying the considered algorithm.
+The convergence theorem can help in solving this problem. Condition \eqref{eq4:reserves} can be considered as a recommendation for the selection of the reserve values, for which the property of
+convergence to the solution of problem \eqref{ex4:problem} remains.
+However, in the general case the values of Lipschitz constants are not given \emph{a priori}, and the length of the interval $[\alpha,\beta]$ from $Q_{m + 1}$, including the sought solution $x^*$ is not kwon also.
+These difficulties are possible to avoid by
+substitution of the unknown coefficients $L_j,\: 1\le j\le m$, by their adaptive underestimates $\mu_\nu$ from (4.15) and by assumption that $\beta-\alpha>\delta=\epsilon q$ where $\epsilon >0$ (accuracy) is a parameter defined in the termination condition and the factor $q>1$ can be interpreted as an a priori estimate of how many times the length of the interval $[\alpha,\beta]$ is greater than given accuracy $\epsilon$.
+As a result, one can take
+the reserve estimates in the form (adaptive reserves)
+\begin{displaymath}
+  \epsilon_\nu=\mu_\nu\delta,\: 1\le\nu\le m.
+\end{displaymath}
+
+\textbf{The results of the experiment.} In order to demonstrate the efficiency of the adaptive reserves, a series of problems with the multiextremal constraints has been solved.
+
+Let us consider a generator $G$ generating the problems with $m$ constraints on the base of a random mechanism $F$ producing the functions $f(x),\: x \in [ a , b ]$. The rules determining the
+functioning of this mechanism $G$ consist in the following.
+\begin{enumerate}
+  \item $m+1$ functions $f_j(x),\: 1\le j \le m+1,\: x\in[a,b]$, are generated independently by means of given mechanism $F$.
+  \item The minimax value
+  \begin{displaymath}
+    \Delta=\min_{a\le x\le b}\max_{1\le j\le m}f_j(x),
+  \end{displaymath}
+  is calculated and the left-hand parts of the constraints
+  \begin{equation}
+    \label{eq4:generated_g}
+    g_j(x)=f_j(x)-\Delta\delta,\: 1\le i\le m,
+  \end{equation}
+  which a nonempty feasible domain $Q$ corresponds to at $\delta>0$ are introduced. Increasing the value of the parameter $\delta$ allows to expand the feasible domain.
+  \item Function \eqref{eq4:generated_g} and the objective function $\varphi(x)=f_{m+1}( x )$ form a problem of the kind \eqref{eq4:problem} obviously having a solution.
+\end{enumerate}
+
+n the cases, when it is necessary to denote the parameters $m$ and $\delta$ used by the generator during its functioning, let us use the notation $G ( m , \delta )$.
+In the experiments, the results of whichare presented below two generators were employed.
+One of these (let us denote it as $G_{SH} ( 5 , 1 ) )$ was based on the function
+\begin{displaymath}
+  \varphi(x)=-\sum_{j=1}^{10}\frac{1}{(K_j(x-A_j)^2+C_j)}x\in[0,10].
+\end{displaymath}
+$1\le K_j\le 3,\: 0\le A_j,\: C_j\le 10 $, where the parameters $K_j,\: A_j,\: C_j ,\: 1 \le j \le 10$ are the independent random variables distributed uniformly over the intervals specified above.
+The second generator hereinafter denoted as $G_{HL} ( 5 , 1 ) )$ was based on the function
+\begin{equation}
+  \varphi(x)=\sum_{j=1}^{14}(A_j\sin(2j\pi x) + B_j\cos(2j\pi x)),\: x\in[0,1]
+\end{equation}
+where the values of the parameters $A_j,\: B_j,\: 1 \le j \le 14$, are distributed independently and uniformly over the interval $[-1 ,1 ]$. Both generators were applied to the generation of 100 own
+problems of the kind \eqref{eq4:problem}.
+
+The problems were solved by the initial index algorithm as well as by the algorithm with $\epsilon$-reservation. The parameters $r_\nu ,\: 1 \le \nu \le 6$, from (4.17) were chosen as
+\begin{displaymath}
+  r_\nu = \left\{
+  \begin{array}{lr}
+    \gamma, & k_\nu < 20\\
+    2, & k_\nu \ge 20
+  \end{array}
+  \right.,
+\end{displaymath}
+where $k_\nu$ is the number of the computed values of the function $g_\nu$ (i.e., the cardinality of the set $I_\nu$).
+In the test set generated by generator $G_{SH}$ the values $\gamma=20,\: \epsilon = 0.001$ and the reserves $\delta_\nu=0.02,\: 1 \le \nu \le 5$, were used in the experiments.
+For the problems from the second test set
+(generated by $G_{HL}$ ) the values $\gamma = 10, \epsilon=0.0001$ and the parameters $\delta_\nu=0.001,\: 1 \le \nu\le 5$, were used.
+
+\begin{figure}[h]
+  \label{fig:4_9}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_9.png}
+  \caption{The operation characteristics in class $G_{SH}$}
+\end{figure}
+
+\begin{figure}[h]
+  \label{fig:4_10}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_10.png}
+  \caption{The operation characteristics in class $G_{HL}$}
+\end{figure}
+
+The operational characteristics for both method obtained for the test problem sets generated by $G_{SH}$ and $G_{HL}$ are presented in Figures 4.9 and 4.10, respectively.
+The lower (solid) curves refer to the initial index algorithm while the upper (dashed) ones correspond to the index algorithm with the adaptive reserves.
+Such arrangement of the curves demonstrates the algorithm with the
+adaptive reserves to provide much faster obtaining the estimates falling into given vicinity of the solution than the initial index algorithm. The average numbers of iterations for the method
+without the reserves and for the method with the adaptive reserves were equal to 171 and 117 for the problems $G_{SH}$ and 199 and 142 for the ones $G_{HL}$ , respectively.
 
@@
 
+Let us now turn to the generalization of the scheme described in sections 3.3 and 3.4 onto the case when the constraint check order is not a fixed one.
+For further consideration, the initial algorithm described above will be referred to as \emph{Fixed Check Order Method} (FCOM).
+
+The algorithm proposed below allows the modification of the constraint order check during solving the problem \eqref{eq4:problem}.
+These modifications are firstly directed onto analyzing during the current iteration the constraint violated at the end of the interval including the current iteration point.
+As it has been already mentioned above, such an approach might reduce the number of checks at the new trial point since the violation of this constraint in a vicinity of already found violation might be quite probable.
+Hence, the need in new rules for the juxtaposition of the indices to the points from \eqref{eq4:search_data} (remind, that these indices determine the use of expressions \eqref{eq4:characteristic} for the computation of the characteristics) as well as in a special order
+of the constraint check at the current trial point arises.
+Note that the formation of such order is inessential for the intervals, the ends of which belong to a feasible domain $Q$ from (4.31) since a necessity for checking all constraints is probable in such intervals.
+
+Assume the constraint with the number $j,\: 1\le j\le m$, to fulfill at all points of the domain
+\begin{equation}
+  Q_j=\{x\in[a,b]:g_j(x)\le 0\},\: 1\le j\le m,
+\end{equation}
+which hereinafter will be called the feasible domain for this constraint. The feasible domain $Q$ of initial problem \eqref{eq4:problem} can be defined by the condition
+\begin{equation}
+  Q=Q_1\cap\dots\cap Q_m.
+\end{equation}
+
+\textbf{The constraints’ check order.} he numbering of the constraints and the objective function introduced in the statement of problem \eqref{eq4:problem} and defined by the sequence of integer numbers
+\begin{equation}
+  \label{eq4:basic_h}
+  H=\{1,2,\dots,m,m+1\},
+\end{equation}
+will be considered as a base one.
+Assume that every particular point $x$ from the search domain $[a,b]$ can have its own order of the problem constraints’ checking.
+The order of such checking corresponding to the point $x_i$ from (4.12) will be defines as some permutation $H(x_i )$ of the basic numbers from \eqref{eq4:basic_h}, i.e.,
+\begin{equation}
+  H(x_i)=\{j_{i1},\dots,j_{im},j_{i,m+1}\},\:0\le i\le k,
+\end{equation}
+where
+\begin{equation}
+  j_{i,m+1}=m+1,\:0\le i\le k.
+\end{equation}
+
+Since the constraints’ check order while the trial execution at the point $x$ i is determined by the corresponding tuple $H(x_i )$ from (4.33), the termination of the trial at this point after the computation of the values of l problem functions produces the value $z_i=g_\nu(x_i)$ where $\nu=j_{il}$.
+In this connection, let us use not the serial number $l$ of the last function calculated in this point, as it was accepted in (4.7) but the index of the last function in the basic numeration (4.32) as the
+value of $\nu(x_i )$ for the point $x_i$ , i.e.,
+\begin{equation}
+  \nu(x_i)=j_{il}.
+\end{equation}
+
+\textbf{Algorithm.} The trial selection rules for Fixed Check Order Method (FCOM) already considered above give the base of the new algorithm scheme, which will be called \emph{Adaptive
+Check Order Method} (ACOM).
+Since the alterable constraint check order excludes a possibility to use expression (4.13) to determine the sets $I\nu$, to construct them, an information array connecting each $x_i$ from (4.12) with the indices of all functions, which have been computed in this node is necessary.
+Hereinafter, the existence of such database is presumed.
+The rest modifications consist in the following.
+
+\emph{Modification 1 (the fixation of the check order)}.
+Each point $x_i$ from \eqref{eq4:search_data} is assigned the corresponding checking order $H(x_i )$ from (4.33), and this order is included into the algorithm database.
+
+\emph{Modification 2 (the determination of the trial points’ indices).} Each point $x_i$ from \eqref{eq4:search_data} is assigned the value $z_i=g_\nu(x_i)$ where according to (4.35), the index $\nu=\nu(x_i)$ is the basic index from (4.32) corresponding to the value of the last function estimated in this point.
+These indices are used in all computations based on expressions (4.13)– (4.17), (4.19).
+
+The modifications in the method of using the indices are related to rules (4.17) of calculations of the characteristics $R(i)$ of the intervals $(x_{i-1},x_i),\: 1\le i\le k$, with noncoincident
+indices of the boundary points only. Assume
+\begin{equation}
+  \nu(x_{i-1})=j_{i-1,p}=j_{i,q}\not=\nu(x_i)=j_{i-1,u}=j_{i,x},
+\end{equation}
+i.e., the problem function with the basic index $\nu(x_{i-1})$ has the index $p$ in the sequence $H(x_{i-1})$ corresponding to the point $x_{i-1}$ and the index $q$ in the sequence $H(x_i)$ corresponding to the point $x_i$.
+Analogously, the function with the basic index $\nu(x_i)$ corresponds to the index $u$ from the sequence $H(x_{i-1})$ and the index s from the sequence $H(x_i)$.
+Then, the introduced modification altering the sixth rule of FCOM consists in the following.
+
+If for the indices $p,q, u$ and $s$ from (4.36) the conditions
+\begin{equation}
+  p<u,q<s,
+\end{equation}
+are fulfilled, in order to determine the characteristic $R(i)$ one should use the second expression from formulae (4.17) with $\nu=\nu(x_i)$ (note that in this case the value $g_\nu(x_{i-1})$ is undefined).
+
+Otherwise, when the reverse to (4.37) conditions are fulfilled
+\begin{equation}
+  p>u,q>s,
+\end{equation}
+the characteristic $R(i)$ is defined by the third expression from (4.17) with $\nu=\nu(x_{i-1})$ (in this
+case, the value $g_\nu(x_i ) is undefined)$.
+In the cases when conditions (4.37) and (4.38) are not fulfilled, the initial order of calculations prescribed by the sixth rule of the initial algorithm remains.
+
+\emph{Modification 3 (the choice of the constraints’ check order in the current trial point).} Assume that the constraints’ check order corresponding to the first two trial points is defined by
+the conditions
+\begin{equation}
+  H(a)=H(b)=H,
+\end{equation}
+where $H$ from (4.32). Let
+\begin{displaymath}
+  p=\nu(x_{t-1}),\: q=\nu(x_t)
+\end{displaymath}
+are the indices of the boundary points of the interval $(x_{t-1},x_t)$ including the current trial point $x^{k+1}$ (according to the eighth rule) determined according to (4.35).
+Let us introduce a notation $\{s,H\backslash s\}$ for the tuple obtained from $H$ by excluding the index $s$, shifting all indices $1,2,\dots,s-1$ to the right of one position, and the transfer of the index s excluded from $H$ into the first position, i.e.,
+\begin{displaymath}
+  \{s,H\backslash s\}=\{s,1,2,\dots,s-1,s+1,\dots,q-1,q+1,\dots,m+1\}.
+\end{displaymath}
+When $1\le p\le q\le m$, the superposition of the operations of the type described above produces the tuple of the kind
+\begin{displaymath}
+  \{p,\{q,H\backslash q\}\backslash p\}=\{p,q,1,\dots,p-1,p+1,\dots,q-1,q+1,\dots,m+1\}.
+\end{displaymath}
+Note that according to (4.34), the index $m+1$ always occupies the last position.
+
+Now let us agree to determine the order $H(x^{k+1}),\: k\ge 1$ according to the following rules:
+\begin{gather}
+ H(x^{k+1})=H,\: p=q=m+1, \nonumber \\
+ H(x^{k+1})=\{p,H\backslash p\}, (p=q<m+1)\cup(p<q=m+1), \nonumber \\
+ H(x^{k+1})=\{q,H\backslash q\}, p=m+1>q\\
+ H(x^{k+1})=\{p, \{q,H\backslash q\}\backslash p\}, p<q\le m \nonumber \\
+ H(x^{k+1})=\{q, \{p,H\backslash p\}\backslash q\} m\ge p>q. \nonumber
+\end{gather}
+
+\textbf{Note.} While solving the problem \eqref{eq4:problem} rules (4.39) and (4.40) included into the scheme of ACOM assign the index values from (4.35) to all trial points unambiguously.
+Then, any feasible point $x$ always corresponds to the index $\nu(x)=m+1$.
+This value can be predicted before the trial execution in such point as well.
+However, if the point $x$ is not a feasible one and does not belong to the points, in which the trials have been executed already, the index value can be undefined in this point.
+
+The convergence conditions for the algorithm are formulated in the following theorem.
+
+\begin{theorem}
+Assume problem \eqref{eq4:problem} to have a solution $x^*$ and the following conditions are fulfilled:
+\begin{enumerate}
+  \item each domain $Q_j ,1\le j\le m+1$ from (4.30) represents a union of a finite number of segments with a positive length;
+  \item each function $g_j(x): 1\le j\le m+1$ satisfies Lipschitz condition with the corresponding constant $L_j$;
+  \item the components $\epsilon_\nu$ of the reserve vector $\epsilon_R$ satisfy the inequalities
+  \begin{equation}
+    0<2\epsilon_\nu<L_\nu(\beta-\alpha),
+  \end{equation}
+  where $\beta-\alpha$ is the length of the interval $[\alpha,\beta]$ belonging to the feasible domain $Q$ from (4.31) and including the point $x^*$, which is the solution of the problem;
+  \item since some sufficiently large step number $k$ the values $\mu_\nu$ from (4.15) corresponding to the nonempty sets $I_\nu$ satisfy the inequalities
+  \begin{displaymath}
+    r_\nu\mu_\nu>2L_\nu,\: 1\le\nu\le m+1,
+  \end{displaymath}
+  where $r_\nu>1,\: 1\le\nu\le m+1$ are the parameters of the algorithm.
+\end{enumerate}
+
+  Then:
+  \begin{enumerate}
+    \item the point $x^*$ is a limit point of the sequence $\{x_k\}$ generated by the described method with the adaptive constraint check order for problem \eqref{eq4:problem} at $\epsilon=0$ in termination condition \eqref{eq4:stop_cond};
+    \item any limit point $x$ of the sequence $\{x_k\}$ is a solution of problem \eqref{eq4:problem};
+    \item the convergence to the limit point $\overline x$ is bilateral if $\overline x\not= a$ and $\overline x\not=b$.
+  \end{enumerate}
+\end{theorem}
+
+\textbf{Proof} of the theorem is given in \cite{}[5].
+
+\textbf{The results of the experiments.} Let us consider solving of the problem from Example \ref{ex4:problem} using ACOM with the parameters $r_\nu=2, \epsilon_\nu=0,\: 1\le \nu\le 3$, and $\epsilon=10^{-5}$ in the termination criterion as an illustration.
+The trial point coordinates executed by the algorithm during the
+problem solving practically coincide to the ones marked by three rows of the vertical strokes in Fig.
+4.6.
+However, the number of the computations of the constraint values appears to be different for FCOM and ACOM.
+In the fixed constraints’ check order the function values have been calculated $k_1 =63,\: k_2 =49,\: k_3 =35$ times while in the adaptive order --- $k_1 =52,\: k_2 =48,\:
+k_3 =35$ times.
+These results are evidence of the decrease of the number of such checks when using ACOM even for the example with two constraints.
+The efficiency of the proposed scheme reveals itself more clearly when solving the problems with many constraints.
+
+With this purpose, let us return to solving the problem series produced by generators $G_{SH}(5,1)$ and $G_{HL} (5,1)$ described in section 3.4. Both generators were applied to the generation of
+its own set of 1000 problems of the kind \eqref{eq4:problem}.
+
+\begin{figure}[h]
+  \label{fig:4_11}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_11.jpg}
+  \caption{The distribution of problems of the class $G_{SH}$}
+\end{figure}
+
+\begin{figure}[h]
+  \label{fig:4_12}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_12.jpg}
+  \caption{The distribution of problems of the class $G_{HL}$}
+\end{figure}
+
+The distribution of the interval lengths $[\alpha,\beta]$ from (4.41) including the solution points $x^*$ corresponding to the problems from the first set is presented in Fig.
+4.11.
+Fig.
+4.12 demonstrates a similar distribution for the problems of the second set.
+In these figures, the values $(\beta=\alpha)/(b-a)$ (in percentage) are plotted against the abscissa axes while the numbers of problems in the corresponding sets are plotted against the ordinate ones.
+In order to obtain the discussed estimates, the lengths of the intervals $[\alpha,\beta]$ corresponding to the respective functions were determined by a numerical method.
+
+More detailed information on the number of problems (in either of two sets) featured by a small (less than 0.1\%) measure of the interval $[\alpha,\beta]$ is presented below.
+
+\begin{table}
+  \caption{The number of problems with a small measure of the interval $[\alpha,\beta]$}
+\begin{center}
+  \label{tab4:exp_results2}
+  \begin{tabular}{| c | c | c | c| c | c | c | c | c| c |}
+    \hline
+    \% & 0.01 & 0.02 & 0.03 & 0.04 & 0.05 & 0.06 & 0.07 & 0.08 & 0.09 \\ \hline
+    $G_{SH}$ & 0 & 1 & 2 & 1 & 0 & 0 & 1 & 1 & 1  \\ \hline
+    $G_{HL}$ & 3 & 6 & 2 & 3 & 5 & 1 & 2 & 1 & 2 \\ \hline
+  \end{tabular}
+\end{center}
+\end{table}
+
+As follows from the figures and Table \ref{tab4:exp_results2}, the problems from the second set produced by generator $G_{HL}$ are characterized in general by a narrower feasible vicinity of the solution point $x^*$.
+
+The constraint functions $g_\nu( x ),\: 1 \le\nu\le 5$, and the objective function $\varphi (x)$ for a problem from the class $G_{HP}$ are plotted in Fig.
+4.13 as an example.
+The feasible domains which are formed by the separate constraints are marked on the abscissa axis.
+The whole feasible domain of the problem (the intersection of all feasible domains of the constraints) is shown along with the graph of the objective function.
+
+\begin{figure}[h]
+  \label{fig:4_13}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_13.png}
+  \caption{Function plots of a test problem from the class $G_{HL}$}
+\end{figure}
+
+The experiments have been carried out for the algorithms described above with the fixed constraints’ check order and with the adaptive one (FCOM and ACOM, respectively).
+The parameters $r_\nu, 1\le\nu\le 5$, from (4.17) were defined by the expressions
+\begin{displaymath}
+  r_\nu = \left\{
+  \begin{array}{lr}
+    \gamma, & k_\nu < 20\\
+    2, & k_\nu \ge 20
+  \end{array}
+  \right.,
+\end{displaymath}
+where $k_\nu$ is the number of the calculated values of the function $g_\nu$ (i.e., the cardinality of the set $I_\nu$).
+In the experiments, the values $\gamma=20,\: \epsilon = 0.001$, and the reserves $\epsilon_\nu=0.1,\: 1\le\nu\le 5$, were set for the set produced by the generator $G_{SH}$.
+For the problems from the second set (generator $G_{HL}$ ) the values  $\gamma=10,\: \epsilon = 0.0001$, and the reserves $\epsilon_\nu=0.001,\: 1\le\nu\le 5$, were used.
+
+\begin{figure}[h]
+  \label{fig:4_14}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_13.png}
+  \caption{Operational characteristics for the class $G_{SH}$}
+\end{figure}
+
+\begin{figure}[h]
+  \label{fig:4_15}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_15.png}
+  \caption{Operational characteristics for the class $G_{HL}$}
+\end{figure}
+
+The operational characteristics for both the methods obtained on the sets produced by generators $G_{SH}$ and $G_{HL}$ are presented in figures 4.14 and 4.15, respectively.
+The lower (solid) curves correspond to FCOM while the upper (dashed) ones --- to ACOM.
+These positions of the curves demonstrate that the algorithm with the adaptive check order provides in average faster obtaining the estimates falling into given vicinity of the solution than the algorithm with the fixed checking order.
+
+\begin{figure}[h]
+  \label{fig:4_16}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_16.png}
+  \caption{Average number of function values computations for the class $G_{SH}$}
+\end{figure}
+
+\begin{figure}[h]
+  \label{fig:4_17}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/4_16.png}
+  \caption{Average number of function values computations for the class $G_{HL}$}
+\end{figure}
+
+Moreover, the trials executed according to the rules of ACOM appear to be much less computation costly (in average) than the ones executed by FCOM.
+This fact is illustrated in figures 4.16 and 4.17, where the average number of computations of the function values executed by FCOM and ACOM for given sets are presented.
+Fig.
+4.16 corresponds to the generator $G_{SH}$ while Fig.
+4.17 shows the results for the generator $G_{HL}$.
+In both figures the function indices are plotted on the abscissa axes.
+The estimates of the average numbers of the computations of these function values during the trials’ execution are marked on the plot axes.
+The solid lines in the figures connect the points of the estimates corresponding to FCOM while the dashed ones connect the points of estimates corresponding to ACOM.
+The presented data confirm that introducing the adaptive constraints’ check order can reduce the total number of the computations essentially.
+
 \section{Parallel computations for the multiextremal optimization problems with nonlinear constraints}
+\subsection{Asynchronous parallel index algorithm}
+In this subsection the asynchronous parallel index algorithm with $p>1$ parallel processors for solving the problem \eqref{ex4:problem} will be presented.
 
@@
 
-\end{document}
+%\end{document}

### 2017-09-24 2017/ParOptBook/Chapter4.tex
parOptBook: update chapter 4

@@
 
-%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\documentclass[graybox]{svmult}
-%
-%% choose options for [] as required from the list
-%% in the Reference Guide
-%
-%\usepackage{mathptmx}       % selects Times Roman as basic font
-%\usepackage{helvet}         % selects Helvetica as sans-serif font
-%\usepackage{courier}        % selects Courier as typewriter font
-%\usepackage{type1cm}        % activate if the above 3 fonts are
-%                           % not available on your system
-%
-%\usepackage{makeidx}         % allows index generation
-%\usepackage{graphicx}        % standard LaTeX graphics tool
-%                             % when including figure files
-%\usepackage{multicol}        % used for the two-column index
-%\usepackage[bottom]{footmisc}% places footnotes at page bottom
-%
-%%custom package
-%\usepackage{mathtools}
-%
-%% see the list of further useful packages
-%% in the Reference Guide
-%
-%\makeindex             % used for the subject index
-%                       % please use the style svind.ist with
-%                       % your makeindex program
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\begin{document}
+% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[graybox]{svmult}
+
+% choose options for [] as required from the list
+% in the Reference Guide
+
+\usepackage{mathptmx}       % selects Times Roman as basic font
+\usepackage{helvet}         % selects Helvetica as sans-serif font
+\usepackage{courier}        % selects Courier as typewriter font
+\usepackage{type1cm}        % activate if the above 3 fonts are
+                           % not available on your system
+
+\usepackage{makeidx}         % allows index generation
+\usepackage{graphicx}        % standard LaTeX graphics tool
+                             % when including figure files
+\usepackage{multicol}        % used for the two-column index
+\usepackage[bottom]{footmisc}% places footnotes at page bottom
+
+%custom package
+\usepackage{mathtools}
+
+% see the list of further useful packages
+% in the Reference Guide
+
+\makeindex             % used for the subject index
+                       % please use the style svind.ist with
+                       % your makeindex program
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{document}
 
@@
 \begin{equation}
-  \label{eq4:mu_const}
+  \label{eq4:z_const}
   z_\nu^\ast = \left\{
@@
 \label{eq4:next_point}
-x^{k+1} = \frac{x_t+x_{t-1}}{2} - \frac{z_t-z_{t-1}}{2r_\nu\mu},\; \nu=\nu(x_{t-1})=\nu(x_t).
+x^{k+1} = \frac{x_t+x_{t-1}}{2} - \frac{z_t-z_{t-1}}{2r_\nu\mu_\nu},\; \nu=\nu(x_{t-1})=\nu(x_t).
 \end{equation}
@@
 \end{equation}
-where t from \eqref{eq4:MaxR} and $\epsilon>0$ is the predefined accuracy of the optimization.
+where $t$ from \eqref{eq4:MaxR} and $\epsilon>0$ is the predefined accuracy of the optimization.
 
@@
 \ref{fig:4_7}, the trial points concentration takes place not only in the vicinity of the point of the global optimum $x^*=2,07957$, but also in the vicinities of the inadmissible boundary points of the domains $Q_j$ from \eqref{eq4:q}.
-This effect can be reduced by the substitution of estimates \eqref{eq4:mu_const} by the estimates
+This effect can be reduced by the substitution of estimates \eqref{eq4:z_const} by the estimates
 \begin{equation}
-  \label{eq4:mu_const}
   z_\nu^\ast = \left\{
@@
 of the constraint check at the current trial point arises.
-Note that the formation of such order is inessential for the intervals, the ends of which belong to a feasible domain $Q$ from (4.31) since a necessity for checking all constraints is probable in such intervals.
+Note that the formation of such order is inessential for the intervals, the ends of which belong to a feasible domain $Q$ from \eqref{eq4:Q_union} since a necessity for checking all constraints is probable in such intervals.
 
@@
 \begin{equation}
+  \label{eq4:Q_union}
   Q=Q_1\cap\dots\cap Q_m.
@@
   \end{equation}
-  where $\beta-\alpha$ is the length of the interval $[\alpha,\beta]$ belonging to the feasible domain $Q$ from (4.31) and including the point $x^*$, which is the solution of the problem;
+  where $\beta-\alpha$ is the length of the interval $[\alpha,\beta]$ belonging to the feasible domain $Q$ from \eqref{eq4:Q_union} and including the point $x^*$, which is the solution of the problem;
   \item since some sufficiently large step number $k$ the values $\mu_\nu$ from (4.15) corresponding to the nonempty sets $I_\nu$ satisfy the inequalities
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{figures/4_16.png}
+  \includegraphics[width=0.8\textwidth]{figures/4_17.png}
   \caption{Average number of function values computations for the class $G_{HL}$}
@@
 In this subsection the asynchronous parallel index algorithm with $p>1$ parallel processors for solving the problem \eqref{ex4:problem} will be presented.
+
+The first $p$ trial points are selected at the internal points of the interval $[a,b]$ arbitrarily.
+Each processor starts the trials execution at the corresponding points.
+After completing the $k^{th}$ trial, when the points $x^{k+1},\dots, x^{k+p - 1}$ have already been selected but the trial results at these points are not obtained yet, the point of the next trial $x^{k+p}$ is selected according to the following rules.
+
+\emph{Rule 1.} Renumber the points of the preceding trials $x^1,\dots,x^k$ by the subscripts in the increasing order of the coordinate values, i.e.,
+\begin{equation}
+  \label{eq4:par_search_data}
+  a=x_0<x_1<\dots x_i<\dots<x_{k+p-1}<x_{k+p}=b,
+\end{equation}
+the points $x_0=a$ and $x_{k + p}=b$ are included into series \eqref{eq4:par_search_data} additionally for the convenience of the future notation.
+
+\emph{Rule 2.} Determine the set $I$ of the points $x_i,\: 1\le i<k+p$, such that the trials at these points have not been completed yet.
+
+\emph{Rule 3.} Juxtapose to each point $x_i,\: 1\le i<k+p$, from series \eqref{eq4:par_search_data} such that $x_i\not\in I$ the
+index $\nu=\nu(x_i)$ and the value $z_i=g_\nu(x_i )$ calculated at this point (the values $z_0$ and $z_{k + p}$ are not determined).
+
+\emph{Rule 4.} Construct the sets
+\begin{equation}
+  I_\nu=\{i:1\le i<k+p, \nu=\nu(x_i),x_i\not\in I\},\:1\le\nu\le m+1,
+\end{equation}
+including the subscripts of all points $x_i ,1\le i<k+p$, with the same index, in which the trials are completed at the moment.
+The boundary points $x_0=a$ and $x_{k + 1}=b$ are considered as the points with the zero indices.
+The additional set $I_0=\{0,k+1\}$ is associated with these points.
+
+Find the maximum value of the index
+\begin{equation}
+  M=\max\{\nu=\nu(x_i),x_i\not\in I,1\le i \le k\}.
+\end{equation}
+
+\emph{Rule 5.} Compute current estimates
+\begin{equation}
+  \label{eq4:lip_est_par}
+  \mu = \max\left\{ \frac{\left|z_i-z_j\right|}{ x_i - x_j }, \; i,j \in I_\nu, \; i>j \right\},
+\end{equation}
+for the unknown Lipschitz constants $L_\nu$ of the functions $g_\nu,\: 1\le\nu\le m+1$.
+If the set $I_\nu$ includes less than two elements or if $\mu_\nu$ from \eqref{eq4:lip_est_par} appears to be equal to zero, to set $\mu_\nu=1$.
+
+\emph{Rule 6.} For all nonempty sets $I_\nu,\: 1\le\nu\le m+1$ calculate the estimates
+\begin{equation}
+  \label{eq4:z_const_par}
+  z_\nu^\ast = \left\{
+  \begin{array}{lr}
+    -\epsilon_\nu, & \nu < M,\\
+    \min\{ g_\nu(y(x_i): i\in I_\nu \}, & \nu = M.
+  \end{array}
+  \right.
+\end{equation}
+where the vector $\epsilon_R=(\epsilon_1 ,\dots, \epsilon_m )$ is a given vector with the non-negative components (\emph{the reserve vector}).
+
+\emph{Rule 7.} Compute the characteristics $R(i)$ for each interval $(x_{i-1}, x_i ),\: x_{i – 1}\not\in I,\: x_i\not\in I, 1\le i\le k+p$ where
+\begin{gather}
+  R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}), \nonumber \\
+  \label{eq4:characteristic_par}
+  R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),\\
+  R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i), \nonumber \\
+  \Delta_i=x_i - x_{i-1} \nonumber
+\end{gather}
+where the values $r_\nu>1,\: 1\le\nu\le m+1$ are the algorithm parameters and $\Delta_i=x_i-x_{i-1}$.
+
+\emph{Rule 8.} Select the interval $(x_{t-1},x_t )$ whith the largest characteristic
+\begin{equation}
+  \label{eq4:MaxR_par}
+  R(t)=\max\{R(i):x_{i-1}\not\in I,x_{i}\not\in I,1\le i\le k+p\}.
+\end{equation}
+
+\emph{Rule 9.} Execute the next trial at the middle point of the interval $(x_{t-1},x_t )$ if the indices of its boundary points are not the same, i.e.,
+\begin{equation}
+  x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
+\end{equation}
+Otherwise, accept
+\begin{equation}
+\label{eq4:next_point_par}
+x^{k+p} = \frac{x_t+x_{t-1}}{2} - \frac{z_t-z_{t-1}}{2r_\nu\mu_\nu},\; \nu=\nu(x_{t-1})=\nu(x_t).
+\end{equation}
+as the point of the next trial.
+
+The rules described above can be supplemented with the termination condition
+\begin{equation}
+\label{eq4:stop_cond_par}
+  x_t - x_{t-1}\le\epsilon
+\end{equation}
+where $t$ from \eqref{eq4:MaxR_par} and $\epsilon>0$ is the predefined accuracy of the search.
+
+\subsection{The convergence conditions of the algorithm}
+\begin{theorem}
+  \label{th4:convergence_par}
+  Assume problem \eqref{eq4:problem} to have a solution $x^*$ and the following conditions to be fulfilled:
+  \begin{itemize}
+    \item each subdomain $Q_j ,1\le j\le m+1$ from \eqref{eq4:q} is a union of a finite number of positive length segments;
+    \item each function $g_j ,1\le j\le m+1$ satisfies Lipschitz condition with the corresponding constant $L_j$;
+    \item the components $\epsilon_\nu$ of the reserve vector $\epsilon_R$ satisfy the inequalities
+    \begin{equation}
+      \label{eq4:reserves_par}
+      0<2\epsilon_\nu<L_\nu(\beta-\alpha)
+    \end{equation}
+    where $\beta-\alpha$ is the length of the interval $[\alpha,\beta]$ belonging to the feasible domain $Q$ from \eqref{eq4:Q_union} and including the point $x^*$ which is the solution of the problem;
+    \item beginning from sufficiently large step number $k$, the values $\nu_\nu$ from \eqref{eq4:lip_est_par} corresponding to the nonempty sets $I_\nu$ satisfy the inequalities
+    \begin{equation}
+      r_\nu\mu_\nu>2L_\nu,\:1\le\nu\le m+1,
+    \end{equation}
+    where the algorithm parameters $r_\nu>1,\:1\le\nu\le m+1$;
+    \item during the parallelization, the inequalities
+    \begin{displaymath}
+      T^j_{l+1}-T^j_j<T<\infty,1\le j\le p, l>1,
+    \end{displaymath}
+    are true where $T_l^j$ is the moment of completing the $l$ th trial by the $j^{th}$ processor and the inequality $1<p<P<\infty$ takes place for the number of processors $p$.
+  \end{itemize}
+  Then:
+  \begin{itemize}
+    \item the point $x^*$ is a limit one for the sequence ${x^k }$ generated by the described asynchronous algorithm for problem \eqref{eq4:problem} if $\epsilon=0$ in termination condition \eqref{eq4:stop_cond_par};
+    \item any limit point $\overline x$ of the sequence $\{x_k \}$ is a solution of the problem \eqref{eq4:problem};
+    \item the convergence to the limit point $\overline x$ is bilateral if $\overline x\not=a$ and $\overline x\not=b$.
+  \end{itemize}
+\end{theorem}
+
+Before the proof of the theorem, let us formulate and prove several auxiliary statements.
+
+\begin{lemma}
+  (on the sequence of the shrinking intervals).
+Assume the conditions of Theorem \ref{th4:convergence_par} to be satisfied.
+Then for each limit point $x$ of the sequence $\{x_k\}$ there exists an infinite sequence of the intervals
+  \begin{equation}
+    \{(x_{t-1},x_t):t=t(q^s)\},\:s=1,2,\dots,
+  \end{equation}
+  satisfying the conditions
+  \begin{gather}
+    \overline x= \cap_{s=1}^\infty[x_{t-1},x_t], \\
+    \lim_{p\to\infty}\Delta_t=0, \\
+    R(t(q^s))>0,\:s=1,2,\dots, \\
+    \lim_{p\to\infty}R(t(q^s))=0,
+  \end{gather}
+  where $q^1<q^2<\dots;\:R(t),\:\Delta_t$, and $t$ are from rules \eqref{eq4:characteristic_par} and \eqref{eq4:MaxR_par}, correspondingly.
+\end{lemma}
+
+\begin{proof}
+  Since $\overline x$ is a limit point, there exists a sequence of the trial numbers $q^1,q^2,\dots$ satisfying the conditions
+  \begin{equation}
+    x^{q+1},\overline x\in[x_{t-1},x_t], t=t(q),q\in\{q^p\},p=1,2,\dots.
+  \end{equation}
+  These conditions reflect the fact that the point of the $(q+1)^{th}$ trial falls into the interval $[x_{t-1},x_t]$ containing the point $\overline x$ at the step $q$, i.e., $t=t(q)$.
+\end{proof}
 

### 2017-09-25 2017/ParOptBook/Chapter4.tex
parOptBook: update chapter 4

@@
 \begin{equation}
+  \label{eq4:lip_condition}
   |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2\in [a,b]
@@
 \begin{equation}
+  \label{eq4:next_point_1_par}
   x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
@@
 \begin{equation}
-\label{eq4:next_point_par}
+\label{eq4:next_point_2_par}
 x^{k+p} = \frac{x_t+x_{t-1}}{2} - \frac{z_t-z_{t-1}}{2r_\nu\mu_\nu},\; \nu=\nu(x_{t-1})=\nu(x_t).
@@
   These conditions reflect the fact that the point of the $(q+1)^{th}$ trial falls into the interval $[x_{t-1},x_t]$ containing the point $\overline x$ at the step $q$, i.e., $t=t(q)$.
+
+  It follows from this condition, from the rules of the next trial point selection \eqref{eq4:next_point_1_par} and \eqref{eq4:next_point_2_par}, and from the estimate of the value $\mu_\nu$ from \eqref{eq4:lip_est_par}, that
+  \begin{equation}
+    \label{eq4:4_60}
+    \max\{x_t-x^{q+1},x^{q+1}-x_{t-1}\}\le\gamma \Delta_t,
+  \end{equation}
+  where $\gamma=(r+1)/2r<1$ and
+  \begin{equation}
+    r=\min\{r_\nu:1\le\nu\le m+1\}>1.
+  \end{equation}
+  Then, the property (4.56) is a consequence of inequality \eqref{eq4:4_60}.
+
+  At every step $k\ge 1$ there exists an interval $(x_{i-1},x_i), \:i=i(k)$, at one of the ends of which the value $z^*_\omega$ from \eqref{eq4:z_const_par} is achieved.
+This value corresponds to the index $\omega$ defined (when $\omega<m$) by the conditions $I_\omega\not=\O,\: I_{\omega+1}\not=\O$.
+If $I_{m+1}=\O$, the index $\omega=m+1$.
+If at that $\nu(x_{i -1} )\not = \nu(x_i )$, then according (4.47) and (4.50), $R(i)=2\Delta_i>0$.
+In the case when $\nu(x_{i -1} )= \nu(x_i )=\omega$, the estimate
+  \begin{displaymath}
+    \max(z_{i-1}, z_i)-z^*_\omega\le\mu\Delta_i
+  \end{displaymath}
+  follows from \eqref{eq4:lip_est_par} and, along with \eqref{eq4:characteristic_par}, gives the estimate
+  \begin{displaymath}
+    R(i)=\Delta_i\left(1-\frac{\max(z_{i-1},z_i)-z^*_\omega}{r_\omega\mu_\omega\Delta_i}\right)\ge (1-r^{-1})\Delta_i,
+  \end{displaymath}
+  where $r$ is from (4.61).
+Thus, at any search step $k\ge 1$ there exists an interval having a strictly positive characteristic.
+Hence, taking into account the rules of selection of the interval with the greatest characteristic \eqref{eq4:MaxR_par}, one can conclude that the intervals from system (4.54) should also have the positive characteristics since these ones undergo splitting up by the trial points falling into these intervals.
+Consequently, conditions (4.57) hold.
+
+  It follows from the trial execution rules and condition (4.35) that for $\nu(x_i )\le m$ the values $z_i$ assigned to the points $x_i$ from \eqref{eq4:par_search_data} are positive.
+Therefore, according to \eqref{eq4:z_const_par},
+  \begin{displaymath}
+    z_i=g_\nu(x_i)\ge z^*_nu,\:1\le\nu=\nu(x_i)\le m+1,\: 0\le i\le k.
+  \end{displaymath}
+    Now from the rules \eqref{eq4:characteristic_par} for the characteristics and the limit (4.56), taking into account the condition \eqref{eq4:lip_condition} of Lipschitzness of the functions and inequality (4.57), one can derive out the truth of statement (4.58).
+\end{proof}
+\begin{lemma}
+  (on the existence of a feasible point). Assume the conditions of Theorem \ref{th4:convergence_par} to be fulfilled. Then, there exists the trial index $h\ge 1$, for which
+  \begin{equation}
+    \label{eq4:62}
+    \nu(x^h)=m+1.
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  Let us assume the opposite and consider the interval $[x_{t-1},x_t]$ including the problem solution $x^*$ from \eqref{eq4:problem} at the step $k$, i.e., $t=t(k)$.
+Since we have assumed the equality \eqref{eq4:62} not to take place and, consequently, the sequence $\{x^k\}$ does not include the points from the feasible domain $Q$, the interval $[\alpha,\beta]$ from (4.52) should satisfy the conditions
+  \begin{equation}
+    x^*\in [\alpha,\beta]\subset[x_{t-1}, x_t].
+  \end{equation}
+  Then, from this condition as well as from the Lipschitz properties of the problem functions the relations
+  \begin{gather}
+    z_{t-1}\le L_\nu(\alpha - x_{t-1})\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t-1}) \\
+    z_t\le L_\nu(x_{t}-\beta)\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t})
+  \end{gather}
+  follow. It was taken into account when deriving out of the above relations that
+  \begin{displaymath}
+    g_\nu(x)\le 0,\: x\in[\alpha,\beta]\subset Q.
+  \end{displaymath}
+  Analogously, the case $\nu=\nu(x_{t-1})=\nu(x_t)$ corresponds to the inequality
+  \begin{equation}
+    z_t+z_{t-1}\le L_\nu \Delta_t - L_\nu(\beta-\alpha).
+  \end{equation}
+
+  Now, from the characteristic rules \eqref{eq4:characteristic_par}, from inequalities (4.64)–(4.66), and from the inequality
+  \begin{displaymath}
+    z^*_\nu\ge-\epsilon_\nu
+  \end{displaymath}
+  resulting from \eqref{eq4:z_const_par} either the estimate
+  \begin{displaymath}
+    R(t)\ge\Delta_t(1-2L_\nu/r_\nu\mu_\nu)+2(L_\nu(\beta-\alpha)-2\epsilon_\nu)/r_\nu\mu_\nu,
+  \end{displaymath}
+  or the estimate
+  \begin{displaymath}
+    R(t)\ge\Delta_t(1-2L_\nu/r_\nu\mu_\nu)+4(L_\nu(\beta-\alpha)-2\epsilon_\nu)/r_\nu\mu_\nu,
+  \end{displaymath}
+  is true. Taking into account the third condition of Theorem \ref{th4:convergence_par} and the fourth one, in both cases we get the inequality
+  \begin{equation}
+    R(t(k))>0
+  \end{equation}
+  which holds for sufficiently large step numbers $k$.
+
+  From (4.58), (4.67), and from the rule (4.48) determining the choice of the interval for the next trial point, the inevitability of carrying out a trial in the interval $[x_{t-1},x_t]$ from (4.63) follows.
+Consequently, the interval $[x_{t-1},x_t]$ will be divided into two subintervals.
+After that, at least one of those will include the point $x^*$.
+The repetition of this division generates (with $k\to\infty$) an infinite sequence of the subintervals shrinking to the point $x^*$ .
+In other words, if the third condition of Theorem \ref{th4:convergence_par} and the fourth one are satisfied, $x^*$ is the limit point of the sequence ${x^k}$.
+
+  Since the coincidence $x^*$ with one of the trial points means (4.62) to hold, then it remains to consider the case when $x^*\in(x_{t- 1} ,x_t ),\: t=t(k)$ for any $k\ge 1$.
+Here, because of the first condition of Theorem \ref{th4:convergence_par} and the definition (4.31) either $[x_{t-1},x^*]\in Q$ or $[x^* ,x_t]\in Q$ if the number $k$ is sufficiently large.
+In this case, either $\nu(x_{t-1})=m+1$ or $\nu (x_t )=m+1.$
+\end{proof}
+
+\begin{lemma}
+  on the feasibility of any limit point). Assume the conditions of Theorem \ref{th4:convergence_par} to be met. Then any limit point $\overline x$ of the sequence $\{x_k\}$ is feasible, i.e.,
+  \begin{equation}
+    \nu(\overline x)=m+1
+  \end{equation}
+\end{lemma}
+
+\begin{proof}
+  Let us assume the opposite. According to (4.55) and (4.56),
 \end{proof}

### 2017-09-28 2017/ParOptBook/Chapter4.tex
parOptBook: update chapter 4

@@
   \begin{equation}
+    \label{eq4:68}
     \nu(\overline x)=m+1
@@
   Let us assume the opposite. According to (4.55) and (4.56),
+  \begin{displaymath}
+    \overline x = \lim_{s\to\infty}x_{t-1}=\lim_{s\to\infty}x_t, t=t(x^s).
+  \end{displaymath}
+  where q s is from (4.57) Therefore, it follows from the assumption $x\in Q$ that there exists such large number $S>1$ and such small subinterval
+  \begin{equation}
+    \label{eq4:69}
+    (u,w)\subset[a,b]\setminus Q,
+  \end{equation}
+  that
+  \begin{equation}
+    [x_{t-1},x_t]\subset(u,w),\:t=t(q^s),\:s\ge S.
+  \end{equation}
+
+  Let us assign to $Q_j, 1\le j\le m$, from (4.30) their open complements $D_j=[a,b]\setminus Q_j$ consisting of finite number of the open intervals with the positive length (as a consequence of the first condition of Theorem \ref{th4:convergence_par}).
+For these complements
+  \begin{equation}
+    g_j(x)>0,x\in D_j,\:1\le j\le m.
+  \end{equation}
+
+  It follows from \eqref{eq4:69} that there exist such subsets of indices $M,N\subset\{1,2,\dots,m\}$ that
+  \begin{equation}
+    (u,\overline x]\subset \cap_{j\in M}D_j,\:[\overline x,w)\subset \cap_{j\in N},
+  \end{equation}
+  where $(u,w)$ is an open interval from \eqref{eq4:69}.
+
+  According to the trial execution rules and to conditions (4.35) and (4.72), for any interval from (4.70) the relations
+  \begin{displaymath}
+    \omega_1=\nu(x_{t-1})\in M,\: \omega_2=\nu(x_t)\in N,\:t=t(q^s),
+  \end{displaymath}
+  hold where $1\le \omega_1,\: \omega_2\le m$. Moreover, according to the rules (4.39) and (4.40) defining the constraints’ check order at trial points,
+  \begin{displaymath}
+    \nu(x_{t-1}),\:\nu(x_t)\in\{\omega_1,\omega2\},\: t=t(q^{p+1}).
+  \end{displaymath}
+  In other words, for sufficiently large number $s>1$ the indices assigned to the boundary points of the considered system of intervals $[x_{t-1},x_t],\: t=t(q^s ): s\ge S$ can take the values from the set $\{\omega_1,\omega2\}$ only.
+
+  Now, from (4.46), (4.74), and (4.57) one can conclude that if the point $x$ is not a feasible one, it cannot be a limit point of the sequence $\{x^k\}$, i.e., for any limit point $x$ the condition \eqref{eq4:68} is true.
+\end{proof}
+
+\begin{proof} (of Theorem \ref{th4:convergence_par}).
+
 \end{proof}

### 2017-09-30 2017/ParOptBook/Chapter4.tex
parOptBook: finish main work on chapter 4

@@
 % RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[graybox]{svmult}
-
-% choose options for [] as required from the list
-% in the Reference Guide
-
-\usepackage{mathptmx}       % selects Times Roman as basic font
-\usepackage{helvet}         % selects Helvetica as sans-serif font
-\usepackage{courier}        % selects Courier as typewriter font
-\usepackage{type1cm}        % activate if the above 3 fonts are
-                           % not available on your system
-
-\usepackage{makeidx}         % allows index generation
-\usepackage{graphicx}        % standard LaTeX graphics tool
-                             % when including figure files
-\usepackage{multicol}        % used for the two-column index
-\usepackage[bottom]{footmisc}% places footnotes at page bottom
-
-%custom package
-\usepackage{mathtools}
-
-% see the list of further useful packages
-% in the Reference Guide
-
-\makeindex             % used for the subject index
-                       % please use the style svind.ist with
-                       % your makeindex program
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
+%\documentclass[graybox]{svmult}
+%
+%% choose options for [] as required from the list
+%% in the Reference Guide
+%
+%\usepackage{mathptmx}       % selects Times Roman as basic font
+%\usepackage{helvet}         % selects Helvetica as sans-serif font
+%\usepackage{courier}        % selects Courier as typewriter font
+%\usepackage{type1cm}        % activate if the above 3 fonts are
+%                           % not available on your system
+%
+%\usepackage{makeidx}         % allows index generation
+%\usepackage{graphicx}        % standard LaTeX graphics tool
+%                             % when including figure files
+%\usepackage{multicol}        % used for the two-column index
+%\usepackage[bottom]{footmisc}% places footnotes at page bottom
+%
+%%custom package
+%\usepackage{mathtools}
+%\usepackage{cite}
+%% see the list of further useful packages
+%% in the Reference Guide
+%
+%\makeindex             % used for the subject index
+%                       % please use the style svind.ist with
+%                       % your makeindex program
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\begin{document}
 
@@
 he exact solution of the problem is $x^*=2.0795,\varphi(x^*)=0.565$. The graphs of the functions $g_1(x),g_2(x)$,and $\varphi(x)$ are presented in Fig. \ref{fig:4_1}.
-\begin{figure}[h]
-  \label{fig:4_1}
+\begin{figure}[ht]
   \centering
@@
   \caption{One-dimensional global optimization problem}
+  \label{fig:4_1}
 \end{figure}
@@
 
-\begin{figure}[h]
-  \label{fig:4_2}
+\begin{figure}[ht]
   \centering
@@
   \caption{A problem with partially computable functions}
+  \label{fig:4_2}
 \end{figure}
@@
 
-\begin{figure}[h]
-  \label{fig:4_3}
+\begin{figure}[ht]
   \centering
@@
   \caption{An ''index'' function}
+  \label{fig:4_3}
 \end{figure}
@@
 Since the set $Q_M$ is always nonempty, problem \eqref{eq4:problem3} always has a solution.
-When $M=m+1$ the solution $x^*=x^*_m+1$  is also the solution of the initial problem \eqref{eq4:problem}.
+When $M=m+1$ the solution $x^*=x^*_{m+1}$  is also the solution of the initial problem \eqref{eq4:problem}.
 When $M<m+1$ the inequality $g_M^*<0$ fulfilled necessarily can be used as an indicator of the incompatibility of the constraints.
@@
 This new function will have discontinuities of the first type at the boundary points of the subdomains $Q_\nu$ from \eqref{eq4:q}.
-Nevertheless, one can estimate the global minimizer $x^*_M$ using the results of $k$ trials for \eqref{eq4:trial} at the points $x_1,\dots, x_k$ in $[a,b]$ (see \cite{3}).
+Nevertheless, one can estimate the global minimizer $x^*_M$ using the results of $k$ trials for \eqref{eq4:trial} at the points $x_1,\dots, x_k$ in $[a,b]$ (see \cite{strongin1992}).
 
@@
 \ref{fig:4_4} is a complement of the set from the right hand side of \eqref{eq4:lip_opt_est} and does not include the optimum point.
-\begin{figure}[h]
-  \label{fig:4_4}
+\begin{figure}[ht]
   \centering
@@
   \caption{An estimate of the optimum point}
+  \label{fig:4_4}
 \end{figure}
@@
 
-This theorem is given without a proof, which can be found, for example, in \cite{}[1, 2].
-Various modifications of this algorithm and the corresponding theory of convergence are given in \cite{}[5–10].
+This theorem is given without a proof, which can be found, for example, in \cite{stronginMarkin1986, stronginMarkin1987}.
+Various modifications of this algorithm and the corresponding theory of convergence are given in \cite{SergeyevFamularoPugliese2001,barkalovStrongin2002,SergeyevFamularoPugliese2003,sergeyev2006,sergeyevKhalafKvasov2007,sergeyevKvasovKhalaf2007}.
 
@@
 
-\begin{figure}[h]
-  \label{fig:4_5}
+\begin{figure}[ht]
   \centering
@@
   \caption{Minimization by means of Penalty Function Method}
+  \label{fig:4_5}
 \end{figure}
@@
 Here $k_i$ is the number of computations of the $i$ problem function, i.e., of the function $g_i(x )$.
-\begin{figure}[h]
-  \label{fig:4_6}
+\begin{figure}[ht]
   \centering
@@
   \caption{The minimization of an ''index'' function}
+  \label{fig:4_6}
 \end{figure}
@@
 }
-\begin{figure}[h]
-  \label{fig:4_7}
+\begin{figure}[ht]
   \centering
@@
   \caption{The solution of a problem by Index Method}
+  \label{fig:4_7}
 \end{figure}
@@
 
-\begin{figure}[h]
-  \label{fig:4_8}
+\begin{figure}[ht]
   \centering
@@
   \caption{The problem solving by Index Method with the $\epsilon$-reservs}
+  \label{fig:4_8}
 \end{figure}
@@
 As it has been demonstrated by the experiment with solving the example described above, the increase of the reserve values may result in an essential decrease of the trial concentration in the boundary points of non-feasible domains.
-The theoretical properties of the trial sequences generated by the index algorithm are determined by the following statement (see the proof in \cite{}[4] also).
+The theoretical properties of the trial sequences generated by the index algorithm are determined by the following statement (see the proof in \cite{stronginSergeyev2000} also).
 \begin{theorem}
@@
 
-\begin{figure}[h]
-  \label{fig:4_9}
+\begin{figure}[ht]
   \centering
@@
   \caption{The operation characteristics in class $G_{SH}$}
-\end{figure}
-
-\begin{figure}[h]
-  \label{fig:4_10}
+  \label{fig:4_9}
+\end{figure}
+
+\begin{figure}[ht]
   \centering
@@
   \caption{The operation characteristics in class $G_{HL}$}
-\end{figure}
-
-The operational characteristics for both method obtained for the test problem sets generated by $G_{SH}$ and $G_{HL}$ are presented in Figures 4.9 and 4.10, respectively.
+  \label{fig:4_10}
+\end{figure}
+
+The operational characteristics for both method obtained for the test problem sets generated by $G_{SH}$ and $G_{HL}$ are presented in Figures \ref{fig:4_9} and \ref{fig:4_10}, respectively.
 The lower (solid) curves refer to the initial index algorithm while the upper (dashed) ones correspond to the index algorithm with the adaptive reserves.
@@
 
-\textbf{Proof} of the theorem is given in \cite{}[5].
+\textbf{Proof} of the theorem is given in \cite{SergeyevFamularoPugliese2001}.
 
@@
 
-\begin{figure}[h]
-  \label{fig:4_11}
+\begin{figure}[ht]
   \centering
@@
   \caption{The distribution of problems of the class $G_{SH}$}
-\end{figure}
-
-\begin{figure}[h]
-  \label{fig:4_12}
+  \label{fig:4_11}
+\end{figure}
+
+\begin{figure}[ht]
   \centering
@@
   \caption{The distribution of problems of the class $G_{HL}$}
+  \label{fig:4_12}
 \end{figure}
@@
 The distribution of the interval lengths $[\alpha,\beta]$ from (4.41) including the solution points $x^*$ corresponding to the problems from the first set is presented in Fig.
-4.11.
+\ref{fig:4_11}.
 Fig.
-4.12 demonstrates a similar distribution for the problems of the second set.
+\ref{fig:4_12} demonstrates a similar distribution for the problems of the second set.
 In these figures, the values $(\beta=\alpha)/(b-a)$ (in percentage) are plotted against the abscissa axes while the numbers of problems in the corresponding sets are plotted against the ordinate ones.
@@
 The constraint functions $g_\nu( x ),\: 1 \le\nu\le 5$, and the objective function $\varphi (x)$ for a problem from the class $G_{HP}$ are plotted in Fig.
-4.13 as an example.
+\ref{fig:4_13} as an example.
 The feasible domains which are formed by the separate constraints are marked on the abscissa axis.
@@
 
-\begin{figure}[h]
-  \label{fig:4_13}
+\begin{figure}[ht]
   \centering
@@
   \caption{Function plots of a test problem from the class $G_{HL}$}
+  \label{fig:4_13}
 \end{figure}
@@
 
-\begin{figure}[h]
-  \label{fig:4_14}
+\begin{figure}[ht]
   \centering
@@
   \caption{Operational characteristics for the class $G_{SH}$}
-\end{figure}
-
-\begin{figure}[h]
-  \label{fig:4_15}
+  \label{fig:4_14}
+\end{figure}
+
+\begin{figure}[ht]
   \centering
@@
   \caption{Operational characteristics for the class $G_{HL}$}
-\end{figure}
-
-The operational characteristics for both the methods obtained on the sets produced by generators $G_{SH}$ and $G_{HL}$ are presented in figures 4.14 and 4.15, respectively.
+  \label{fig:4_15}
+\end{figure}
+
+The operational characteristics for both the methods obtained on the sets produced by generators $G_{SH}$ and $G_{HL}$ are presented in figures \ref{fig:4_14} and \ref{fig:4_15}, respectively.
 The lower (solid) curves correspond to FCOM while the upper (dashed) ones --- to ACOM.
@@
 
-\begin{figure}[h]
-  \label{fig:4_16}
+\begin{figure}[ht]
   \centering
@@
   \caption{Average number of function values computations for the class $G_{SH}$}
-\end{figure}
-
-\begin{figure}[h]
-  \label{fig:4_17}
+  \label{fig:4_16}
+\end{figure}
+
+\begin{figure}[ht]
   \centering
@@
   \caption{Average number of function values computations for the class $G_{HL}$}
+  \label{fig:4_17}
 \end{figure}
@@
 \begin{lemma}
+  \label{lm4:2}
   (on the existence of a feasible point). Assume the conditions of Theorem \ref{th4:convergence_par} to be fulfilled. Then, there exists the trial index $h\ge 1$, for which
@@
   \begin{equation}
+    \label{eq4:63}
     x^*\in [\alpha,\beta]\subset[x_{t-1}, x_t].
@@
   \begin{equation}
+    \label{eq4:67}
     R(t(k))>0
@@
 
-  From (4.58), (4.67), and from the rule (4.48) determining the choice of the interval for the next trial point, the inevitability of carrying out a trial in the interval $[x_{t-1},x_t]$ from (4.63) follows.
+  From (4.58), \eqref{eq4:67}, and from the rule (4.48) determining the choice of the interval for the next trial point, the inevitability of carrying out a trial in the interval $[x_{t-1},x_t]$ from (4.63) follows.
 Consequently, the interval $[x_{t-1},x_t]$ will be divided into two subintervals.
@@
 \begin{lemma}
+  \label{lm4:3}
   on the feasibility of any limit point). Assume the conditions of Theorem \ref{th4:convergence_par} to be met. Then any limit point $\overline x$ of the sequence $\{x_k\}$ is feasible, i.e.,
@@
   \begin{displaymath}
-    \nu(x_{t-1}),\:\nu(x_t)\in\{\omega_1,\omega2\},\: t=t(q^{p+1}).
-  \end{displaymath}
-  In other words, for sufficiently large number $s>1$ the indices assigned to the boundary points of the considered system of intervals $[x_{t-1},x_t],\: t=t(q^s ): s\ge S$ can take the values from the set $\{\omega_1,\omega2\}$ only.
+    \nu(x_{t-1}),\:\nu(x_t)\in\{\omega_1,\omega_2\},\: t=t(q^{p+1}).
+  \end{displaymath}
+  In other words, for sufficiently large number $s>1$ the indices assigned to the boundary points of the considered system of intervals $[x_{t-1},x_t],\: t=t(q^s ): s\ge S$ can take the values from the set $\{\omega_1,\omega_2\}$ only.
 
@@
 
+  I.
+Let us prove the first assertion of Theorem \ref{th4:convergence_par} based on Lemma \ref{lm4:2} and Lemma \ref{lm4:3}, i.e., that the point $x^*$ from \eqref{eq4:problem} is the limit point.
+Let us assume the opposite and consider two  possible cases.
+Assume $[x_{t-1},x_t],\: t=t(k)$ to be an interval including the point $x^*$ at the step $k$.
+Consider the condition
+  \begin{equation}
+    \label{eq4:75}
+    \max\{\nu(x_{t-1}),\nu(x_t)\}=m+1
+  \end{equation}
+  for the indices of the boundary points of the given interval and assume that this condition is not satisfied for sufficiently large $k$.
+In this case, inclusions \eqref{eq4:63} should take place.
+This situation has been already considered in Lemma \ref{lm4:2}, where its incompatibility with the assumption of the point $x^*$ is not the limit point has been demonstrated.
+
+  So, it remains to consider the case when condition \eqref{eq4:75} is fulfilled for some subsequence of the step numbers $k$. Then, the relations
+  \begin{gather}
+    z_{t-1}\le g_\nu(x^*)+L_\nu(x^*-x_{t-1}),\: \nu=\nu(x_{t-1}), \nonumber \\
+    z_t\le g_\nu(x^*)+L_\nu(x_t-x^*),\: \nu=\nu(x_t), \nonumber
+  \end{gather}
+  follow from \eqref{eq4:lip_condition} and the equality $g_{m+1}(x^*)=\varphi(x^*)$ and the inequality
+  \begin{displaymath}
+    z_t+z_{t-1}\le 2\varphi(x^*)+L_{m+1}\Delta_t,
+  \end{displaymath}
+  correspond to the case
+  \begin{equation}
+    \label{eq4:76}
+    \nu(x_{t-1})=\nu(x_t)=m+1,
+  \end{equation}
+
+  From (4.47) taking into account \eqref{eq4:75} and the inequality $z_{m + 1}\ge \varphi(x)$ resulting from (4.46), one can derive the estimate
+  \begin{equation}
+    \label{eq4:77}
+    R(t)\ge \gamma\Delta_t(1-2L_{m+1}/(r_{m+1}\mu_{m+1}))
+  \end{equation}
+  where the coefficient $\gamma=1$ in the case when equality \eqref{eq4:76} holds. Otherwise, $\gamma= 2$.
+
+  Now, it follows from \eqref{eq4:77} and from the fourth condition of Theorem \ref{th4:convergence_par} that inequality \eqref{eq4:67} holds for some subsequence of $k$ that, as it has been already discussed above, leads to the contradiction to the assumption of the non-limit character of the point $x^*$.
+
+  II.
+Let us demonstrate that any limit point $\overline x$ of the sequence $\{x^k\}$ generated by ACOM is a solution of the problem \eqref{eq4:problem} under the conditions of the theorem.
+Since the point $x^*$ from \eqref{eq4:problem} is a limit one, then, according to (4.46),
+  \begin{equation}
+    \label{eq4:78}
+    \lim_{p\to\infty}z^*_{m+1}=\varphi(x^*).
+  \end{equation}
+
+  Let $x\not=x^*$ is another limit point of the sequence $\{x^k\}$ being a feasible one according to \eqref{eq4:68}.
+Then, for the characteristic of the interval $[x_{j-1},x_j]$ including the point $\overline x$ at the step $k$, because of Lipschitz condition \eqref{eq4:lip_condition}, (4.45), (4.47), (4.56), and \eqref{eq4:78} the relation
+  \begin{displaymath}
+    \lim_{k\to\infty}R(j(k))\le-4(\varphi(\overline x)-\varphi(x^*))/(r_{m+1}L_{m+1}).
+  \end{displaymath}
+  holds.
+
+  If $\varphi(x)>\varphi(x^*)$ the characteristic of the interval $[x_{j-1},x_j]$ including the point $\overline x$ will become negative starting from some step that contradicts to property (4.57), which should be fulfilled for any limit point.
+Consequently, the equality $\varphi(x)=\varphi(x^*)$  is true.
+
+  Theorem is proved.
 \end{proof}
@@
 For all proposed algorithms, the results of numerical experiments are presented. The results confirm the theoretical statements about convergence acceleration.
-
-\end{document}
+\begin{thebibliography}{99.}
+  \bibitem{stronginMarkin1986} Strongin, R.G., Markin, D.L.: Minimization of multiextremal functions under nonconvex constraints. Cybernetics \textbf{22}(4), 486---493 (1986)
+  \bibitem{stronginMarkin1987} Markin, D.
+L., Strongin, R.
+G.: A Method for Solving Multi-Extremal Problems with Non-Convex Constraints, That Uses a Priori Information about Estimates of the Optimum.
+Comput.
+Maths.
+Math.
+Phys.
+\textbf{27}(1), 33---39 (1987)
+  \bibitem{strongin1992} Strongin, R.G.: Algorithms for multi-extremal mathematical programming problems employing the set of joint space-filling curves.
+Journal of Global Optimization \textbf{2}(4), 357--378 (1992)
+  \bibitem{stronginSergeyev2000} Strongin, R.G., Sergeyev, Y.D.: AGlobal Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000)
+  \bibitem{SergeyevFamularoPugliese2001} Sergeyev Ya.D., Famularo D., Pugliese P.: Index Branch-and-Bound Algorithm for Lipschitz univariate global optimization with multiextremal constraints.
+Journal of Global Optimization \textbf{21}(3), 317---341 (2001)
+  \bibitem{barkalovStrongin2002} Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints.
+Comput.
+Maths.
+Math.
+Phys.
+\textbf{42}(9), 1289---1300 (2002)
+  \bibitem{SergeyevFamularoPugliese2003} Sergeyev Ya.D., Pugliese P., Famularo D.: Index information algorithm with local tuning for solving multidimensional global optimization problems with multiextremal constraints.
+Mathematical Programming \textbf{96}(3), 489---512.
+(2003)
+  \bibitem{sergeyev2006} Sergeyev Ya.D.: Univariate global optimization with multiextremal non-differentiable constraints without penalty functions.
+Computational Optimization and Applications \textbf{34}(2), 229---248 (2006)
+  \bibitem{sergeyevKhalafKvasov2007} Sergeyev, Y.D., Khalaf, F.M.H., Kvasov, D.E.: Univariate algorithms for solving global optimization problems with multiextremal non-differentiable constraints.
+In: A.
+T\"orn, J.
+ä ilinskas (eds.) Models and Algorithms for Global Optimization, Springer, 123---140 (2007)
+  \bibitem{sergeyevKvasovKhalaf2007} Sergeyev Ya.D., Kvasov D.E., Khalaf F.M.H.: A one-dimensional local tuning algorithm for solving GO problems with partially defined constraints.
+Optimization Letters \textbf{1}(1), 85---99 (2007)
+\end{thebibliography}
+%\end{document}

### 2017-10-03 2017/ParOptBook/Chapter4.tex
Chapter 6 is ready

@@
   \label{eq4:problem}
-  \varphi(x^*)=\min\{\varphi(x):x\in [a,b], g_j(x)\le 0,1\le j\le m\}
+  \varphi(x^*)=\min\{\varphi(x):x\in [a,b], g_j(x)\le 0,\; 1\le j\le m\}
 \end{equation}
@@
 \begin{table}
-  \caption{}
+  \caption{The results of solving the test problem}
 \begin{center}
   \label{tab4:exp_results}
-  \begin{tabular}{| l | c | c | c| }
+  \begin{tabular}{cccc}
     \hline
-    & $k_1$ & $k_2$ & $k_3$ \\ \hline
-    Scanning a uniform grid & 160000 & 90280 & 56476 \\ \hline
-    Penalty Function Method &  375 & 375 & 375 \\ \hline
+    Method & $k_1$ & $k_2$ & $k_3$ \\ \hline
+    Uniform grid technique  & 160000 & 90280 & 56476 \\ 
+    Penalty Function Method &  375 & 375 & 375 \\ 
     Index method & 63 & 49 & 35 \\ \hline
@@
   \includegraphics[width=0.8\textwidth]{figures/4_6.png}
-  \caption{The minimization of an ''index'' function}
+  \caption{The minimization of an ``index'' function}
   \label{fig:4_6}
@@
 \begin{displaymath}
-  \varphi(x^*)=\min\{\varphi(x):x\in [a,b], g_j(x)\le 0,1\le j\le m\}
+  \varphi(x^*)=\min\{\varphi(x):x\in [a,b],\; g_j(x)\le 0, \; 1\le j\le m\}
 \end{displaymath}
@@
 {
-Let us consider problem \eqref{eq4:problem} with $x\in[0,6,2,2]$, $m=3$ as an illustration:
+Let us consider problem \eqref{eq4:problem} with $x\in[0.6,2.2]$, $m=3$ as an illustration:
 \begin{gather*}
@@
 \section{Index method with adaptive order of checking for constraints}
-According to the rules of the index method considered in the preceding subsections, every trial includes the successive check of the fulfillment of the problem constraints in the corresponding point of the search domain.
+According to the rules of the index method considered in the preceding subsections, every trial includes the successive check of the fulfilment of the problem constraints in the corresponding point of the search domain.
 The finding of the first violated constraint terminates the trial and initiates the transition to the next iteration.
@@
 
-For many applied problems, it is naturally to assume that a constraint violated at a point in the search domain, will be violated in some neighborhood of this point as well.
-In this case, it might be useful to begin the next trials in the points of this neighborhood with checking the fulfillment of this particular constraint.
+For many applied problems, it is naturally to assume that a constraint violated at a point in the search domain, will be violated in some neighbourhood of this point as well.
+In this case, it might be useful to begin the next trials in the points of this neighbourhood with checking the fulfilment of this particular constraint.
 In the scope of the above, it seems reasonable to design a global optimization algorithm allowing the modification of the constraint  check order taking into account the information obtained during the search.
@@
 
-Assume the constraint with the number $j,\: 1\le j\le m$, to fulfill at all points of the domain
+Assume the constraint with the number $j,\: 1\le j\le m$, is fulfilled at all points of the domain
 \begin{equation}
@@
   \label{tab4:exp_results2}
-  \begin{tabular}{| c | c | c | c| c | c | c | c | c| c |}
+  \begin{tabular}{cccccccccc}
     \hline
-    \% & 0.01 & 0.02 & 0.03 & 0.04 & 0.05 & 0.06 & 0.07 & 0.08 & 0.09 \\ \hline
-    $G_{SH}$ & 0 & 1 & 2 & 1 & 0 & 0 & 1 & 1 & 1  \\ \hline
+     & 0.01\% & 0.02\% & 0.03\% & 0.04\% & 0.05\% & 0.06\% & 0.07\% & 0.08\% & 0.09\% \\ \hline
+    $G_{SH}$ & 0 & 1 & 2 & 1 & 0 & 0 & 1 & 1 & 1  \\ 
     $G_{HL}$ & 3 & 6 & 2 & 3 & 5 & 1 & 2 & 1 & 2 \\ \hline
@@
 Each processor starts the trials execution at the corresponding points.
-After completing the $k^{th}$ trial, when the points $x^{k+1},\dots, x^{k+p - 1}$ have already been selected but the trial results at these points are not obtained yet, the point of the next trial $x^{k+p}$ is selected according to the following rules.
+After completing the $k$-th trial, when the points $x^{k+1},\dots, x^{k+p - 1}$ have already been selected but the trial results at these points are not obtained yet, the point of the next trial $x^{k+p}$ is selected according to the following rules.
 
@@
     \end{displaymath}
-    are true where $T_l^j$ is the moment of completing the $l$ th trial by the $j^{th}$ processor and the inequality $1<p<P<\infty$ takes place for the number of processors $p$.
+    are true where $T_l^j$ is the moment of completing the $l$-th trial by the $j$-th processor and the inequality $1<p<P<\infty$ takes place for the number of processors $p$.
   \end{itemize}
@@
   \end{equation}
-  These conditions reflect the fact that the point of the $(q+1)^{th}$ trial falls into the interval $[x_{t-1},x_t]$ containing the point $\overline x$ at the step $q$, i.e., $t=t(q)$.
+  These conditions reflect the fact that the point of the $(q+1)$-th trial falls into the interval $[x_{t-1},x_t]$ containing the point $\overline x$ at the step $q$, i.e., $t=t(q)$.
 
@@
   At every step $k\ge 1$ there exists an interval $(x_{i-1},x_i), \:i=i(k)$, at one of the ends of which the value $z^*_\omega$ from \eqref{eq4:z_const_par} is achieved.
-This value corresponds to the index $\omega$ defined (when $\omega<m$) by the conditions $I_\omega\not=\O,\: I_{\omega+1}\not=\O$.
-If $I_{m+1}=\O$, the index $\omega=m+1$.
+This value corresponds to the index $\omega$ defined (when $\omega<m$) by the conditions $I_\omega\not= \varnothing,\: I_{\omega+1} =  \varnothing$.
+If $I_{m+1}= \varnothing$, the index $\omega=m+1$.
 If at that $\nu(x_{i -1} )\not = \nu(x_i )$, then according (4.47) and (4.50), $R(i)=2\Delta_i>0$.

### 2017-10-03 2017/ParOptBook/Chapter6.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter4.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter5.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter6.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter7.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter8.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\dedic.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\editor.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\foreword.tex
D:\Common\Barkalov\Публикации\20

--- 2017-09-01 2017/ParOptBook/Chapter7.tex
@@
 %\usepackage[bottom]{footmisc}% places footnotes at page bottom
-%
+%\usepackage{amsmath}
+
+%\usepackage[colorlinks=true]{hyperref}
+%\hypersetup{urlcolor=blue, citecolor=red}
+
 %% see the list of further useful packages
@@
 
-\title{Software Systems for Global Optimization }
-% Use \titlerunning{Short Title} for an abbreviated version of
-% your contribution title if the original one is too long
-\author{Name of First Author and Name of Second Author}
-% Use \authorrunning{Short Title} for an abbreviated version of
-% your contribution title if the original one is too long
-\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
-\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
+\title{Parallel computations for the multidimensional multiextremal optimization problems with the dimensionality reduction using Peano curves }
+\titlerunning{Parallel computations for the multidimensional optimization}
+
+\author{Konstantin Barkalov, Victor Gergel, Roman Strongin}
+\authorrunning{K. Barkalov, V.Gergel, R. Strongin}
+
+\institute{Konstantin Barkalov,  Victor Gergel, Roman Strongin \at Lobachevsky State University of Nizhni Novgorod,  Nizhni Novgorod, Russia \email{konstantin.barkalov@itmm.unn.ru}}
 %
@@
 
-\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\abstract*{The chapter considers dimensionality reduction scheme based on Peano-type space-filling cures (evolvents).
+The initial multidimensional optimization problem is substituted by an equivalent one-dimensional one based on the application of the mapping of a multidimensional search domain onto a segment of the real axis.
+A general description of the approach and its substantiations are given.
+The development of the classical scheme based on the application of a set of mappings and the implementation of new types of evolvents is considered for the parallel algorithm.}
 
-\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\abstract{The chapter considers dimensionality reduction scheme based on Peano-type space-filling cures (evolvents).
+The initial multidimensional optimization problem is substituted by an equivalent one-dimensional one based on the application of the mapping of a multidimensional search domain onto a segment of the real axis.
+A general description of the approach and its substantiations are given.
+The development of the classical scheme based on the application of a set of mappings and the implementation of new types of evolvents is considered for the parallel algorithm.}
 
-\section{Section Heading}
-Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.
+\section{General approach scheme}
+
+Let us consider a multidimensional global optimization problem
+\begin{equation}\label{6_problem} 
+\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_j(y)\leq 0, \; 1 \leq j \leq m\right\}},
+\end{equation} 
+\begin{equation}\label{6_D}
+D=\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1\leq i \leq N \right\}.
+\end{equation}
+This problem statement covers a large class of problems since any hyperinterval 
+\[
+S=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}
+\]
+can be reduced to the  hypercube (\ref{6_D}) by linear transformation of coordinates.
+
+There is a number of ways to adapt efficient one-dimensional algorithms for solving multidimensional problems; see, for example, the diagonal partitions method in \cite{6_Sergeyev2006,6_Sergeyev2015,6_Sergeyev2017} or the simplicial partitions method in \cite{6_Zilinskas2008,6_Zilinskas2014,6_Zilinskas2014_1}.
+
+The dimensionality reduction method described below uses a mapping of a multidimensional search domain onto a one-dimensional interval by means of so called \textit{Peano curves}.
+Peano, an Italian mathematician, has found that an single-valued continuous mapping $y(x)=(y_1(x),y_2(x))$ of the interval $[0,1]$ onto the square
+\[
+\left\{y\in R^2: -2^{-1}\leq y_1,y_2 \leq 2^{-1} \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\}
+\]
+can be constructed.
+This result was generalized onto a multidimensional case, i.e., the existence of the curves $y(x)$ defined by continuous coordinate functions $y_i(x), \; x \in [0,1],\; 1 \leq i \leq N,$ and mapping uniquely the unit interval $[0,1]$ onto a $N$-dimensional hypercube
+\[
+D=\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1\leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\}
+\]
+had been proven.
+These curves, called also \textit{Peano curves}, allow to reduce a multidimensional constrained optimization problem over the domain $D$ to a one-dimensional constrained minimization problem over the unit interval $[0,1]$
+\begin{equation}\label{6_problem_1} 
+\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)):x\in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}}.
+\end{equation}
+
+The considered dimensionality reduction scheme juxtaposes to a multidimensional problem with Lipschitz objective function and Lipschitz constraints a one-dimensional problem, where the corresponding functions satisfy uniform H\"{o}lder condition (see \cite{6_Strongin2000,6_Strongin2013}), i.e.,
+\begin{equation}\label{6_Holder}
+\left|g_j(y(x'))-g_j(y(x''))\right| \leq K_j \left|x'-x''\right|^{1/N}, \; x',x''\in [0,1], \; 1\leq j \leq m+1. 
+\end{equation}
+Here $N$ is the dimensionality of the initial multidimensional problem and the coefficients $K_j$ are related with Lipschitz constant $L_j$ of the initial problem as 
+\[
+K_j \leq 2L_j \sqrt{N+3},\; 1\leq j \leq m+1.
+\]
+Some issues of H\"older functions optimization are considered in \cite{6_Gourdin,6_Lera2002,6_Lera2010,6_Hime,6_Lera2015}.
+
+We will consider initial problem (\ref{6_problem}) assuming that the problem functions $g_j(y),\; 1 \leq j \leq m+1,$ may be defined partially.
+This means that they are defined and computable only in the subranges $Q_j \in [0,1]$,
+\begin{equation}\label{6_Q}
+Q_1=[0,1], \; Q_{j+1}=\left\{x \in Q_j : g_j(y(x)) \leq 0 \right\}, \; 1 \leq j \leq m.
+\end{equation}
+These conditions allows to introduce a classification of the points $x \in [0,1]$ according to the number $\nu = \nu(x)$ of the constraints computed at this point.
+The index $\nu(x)$ can also be defined by the conditions
+\begin{equation}\label{6_nu}
+g_j(y(x)) \leq 0, \; 1 \leq j < \nu, \; g_\nu(y(x))>0,
+\end{equation}
+where the last inequality is inessential if $\nu=m+1$.
+
+Therefore, the execution of an iteration at the point $x \in [0,1]$ is reduced to the successive computation of the values $g_j(y(x)),\; 1 \leq j \leq \nu(x),$ i.e., the next value $g_{j+1}(y(x))$ is computed if and only if $g_j(y(x)) \leq 0$.
+The process of computations is terminated either as a result of occurrence of the inequality $g_j(y(x))>0$ or by satisfying the equality $\nu(x)=m+1$.
+Moreover, the described procedure called the trial at the point $x \in [0,1]$ leads automatically to determining the index $\nu$ of this point.
+%since conditions (\ref{6_nu}) are equivalent to the inequalities
+The pair of values 
+\[
+\nu=\nu(x), \; z=g_\nu(y(x)),
+\]
+generated by the trial at the point $x \in [0,1]$, is called \textit{the trial result}.
+
+Besides the exact solution $y^\ast$ from (\ref{6_problem}) let us consider also the \textit{$\epsilon$-reserved solution} of the problem (\ref{6_problem}) defined by the conditions 
+\begin{equation}\label{6_eps_r}
+\varphi(y_\epsilon) = \min \left\{ \varphi(y): \; y \in D, \; g_j(y) \leq -\epsilon_j, \; 1 \leq j \leq m \right\},
+\end{equation}
+where $\epsilon_1,...,\epsilon_m$ are the positive values (the ``reserves'' with respect to each constraint). Let us introduce into the consideration also the set
+\begin{equation}\label{6_eps_solution}
+Y_\epsilon =  \left\{ y \in D: \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_\epsilon) \right\}
+\end{equation}
+of all feasible points of problem (\ref{6_problem}), which are not worse (with respect to the objective function value) than the $\epsilon$-reserved solution.
+Fig.~\ref{6_fig_1} illustrates the definition introduced.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.7\linewidth]{figures/6_1.png}
+\caption{$\epsilon$-reserved solution}
+\label{6_fig_1}     
+\end{figure}
+
+The existence of the $\epsilon$-reserved solution of problem (\ref{6_problem}) can be interpreted as an analog of the regularity conditions in the classical problems of nonlinear programming.
+It should be noted also the applied importance of this condition.
+Even if the exact solution $y^\ast$ is known, its practical realization is possible as some approximation of $y^\ast$ only.
+Therefore, it is important that there exist the points from the feasible domain $Q$ close to $y^\ast$ (with respect to the coordinates and values).
+The existence of the $\epsilon$-reserved solution guarantees the existence of such points.
+In this case, the domain $Y_\epsilon$ can play a role of the set of suitable approximations.
+
+Prior to going to the detailed description of the Peano curves and of the algorithm, let us consider some properties of reduced problem (\ref{6_problem_1}), which are absent in the pure one-dimensional case.
+First, as it has been already mentioned above, the functions $g_j(y(x)), \; 1 \leq j \leq m+1,$ are not Lipschitzian.
+They satisfy H\"older condition.
+Consequently, all the interval lengths in the search algorithm rules have to be replaced by the lengths in the metric
+\[
+\rho (x',x'') = \left(\left|x'-x''\right|\right)^{1/N}, \; x',x'' \in [0,1].
+\]
+Second, in general, sets (\ref{6_Q}) being the inverse images of the essentially non-cubic $N$-dimensional subsets with the nonlinear boundaries $g_j(y)=0, \; y \in D, \; 1\leq j \leq m,$ are not the finite unions of $x$-intercepts.
+Consequently, taking into account the $\epsilon$-reserved solutions plays a significantly more important role for the convergence in the reduced univariate problem (\ref{6_problem_1}) than in the pure one-dimensional case.
+
+The issues of the numerical construction of the Peano-type space filling curves and the corresponding theory are considered in details in \cite{6_Butz,6_Sagan,6_Strongin2000,6_Strongin2013}.
+Here we will describe the basic steps of constructing a curve mapping a unit interval of the real axis $[0,1]$ onto a hypercube
+\[
+D=\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1\leq i \leq N \right\}.
+\]
+
+\begin{enumerate}
+	\item 
+A hypercube $D$ with the unit edge length is divided by the coordinate hyperplanes into $2^N$ hypercubes of the first partition (with the edge length equal to $1/2$), which are indexed by the numbers $z_1$ ranging from $0$ to $2^N-1$.
+Let us agree to designate the hypercube of the first partition with the index $z_1$ as $D(z_1)$.
+
+Next, each hypercube of the first partition is also divided into $2^N$ hypercubes of the second partition (with the edge length equal to $1/4$) by the hyperplanes parallel to the coordinate axes and crossing the middles of the hypercube edges orthogonal to these hyperplanes.
+The hypercubes of the second partition  included in the hypercube $D(z_1)$ are indexed by the numbers $z_2$ ranging from $0$ to $2^N-1$.
+The hypercube of the second partition  with the index $z_2$ from  $D(z_1)$ is denoted as $D(z_1, z_2)$.
+The case of $N=2$ is presented in Fig.~\ref{6_fig_2} for $m=1$, $m=2$.
+
+\begin{figure}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=0.9\linewidth]{figures/6_2a.png} \\ $m=1$}
+\end{minipage}
+\hfill
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=0.9\linewidth]{figures/6_2b.png} \\ $m=2$}
+\end{minipage}
+\caption{The partition of an $N$-dimensional hypercube}
+\label{6_fig_2}
+\end{figure}
+
+Continuing this process, one can construct the hypercubes of any $m$-th partition with the edge length equal to $(1/2)^m$, which are denoted as $D(z_1,...,z_m)$.
+It is obvious, that $D(z_1) \supset D(z_1,z_2) \supset ...
+\supset D(z_1,...,z_m)$ and $0 \leq z_j \leq 2^N-1$, \mbox{$1 \leq j \leq m$}.
+
+\item
+Now let us perform the division of the interval $[0,1]$ into $2^N$ equal parts, each of them also is divided into $2^N$ equal parts, etc.
+The elements of each partition are indexed from the left to the right by the numbers $z_j$ ranging from $0$ to \mbox{$2^N-1$} where $j$ is the index of subdivision.
+Let us denote the intervals of the $m$-th partition as $d(z_1,...,z_m)$ where, for example, $d(z_1,z_2)$ means the interval of the second partition with the index $z_2$ being a part of the interval $d(z_1)$ of the first partition with the index $z_1$.
+
+One can note that  $d(z_1) \supset d(z_1,z_2) \supset ...
+\supset d(z_1,...,z_m)$ and the length of the interval $d(z_1,...,z_m)$ is equal to  $(1/2)^{mN}$.
+An interval $d(z_1,...,z_m)$ is supposed to include its left hand end.
+It contains its right hand boundary if and only if when $z_1=z_2=...=z_m=2^N-1$.
+The case of $N=2$ is presented in Fig.~\ref{6_fig_3} for $m=1$ and $m=2$.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.9\linewidth]{figures/6_3.png}
+\caption{Partition of a one-dimensional interval}
+\label{6_fig_3}     
+\end{figure}
+
+\item
+Assume that the point $y(x) \in D$ corresponding to the point $x \in [0,1]$ for any $m \geq $ belongs to the hypercube $D(z_1,...,z_m)$ if $x$ falls into the interval $d(z_1,...,z_m)$, i.e.,
+\[
+x \in d(z_1,...,z_m) \rightarrow y(x) \in D(z_1,...,z_m).
+\]
+The constructed correspondence  $y(x)$ is single-valued.
+\item
+In order to provide the continuity of the constructed correspondence, let us impose the following requirements on the indexing order of the hypercubes of each partition.
+Since $2^{mN}$ centers $y(z_1,...,z_m)$ of the hypercubes of the $m$-th partition $D(z_1,...,z_m)$ form a uniform orthogonal grid in the domain $D$ (the step of this grid with respect to any coordinate is equal to $2^{-m}$), one can introduce the following numeration of the grid nodes.
+Let us enumerate from the left to the right by index $i$ all subintervals of the $m$- 	th partition in the interval $[0,1]$, i.e.,
+\[
+d(z_1,...,z_m)=[x_i, x_{i+1}), \; 0 \leq i < 2^{mN}-1,
+\]
+were the left end of the $i$-th interval is denoted as $x_i$.
+
+Assume the center of the hypercube $D(z_1,...,z_m)$ to have the same index $i$ as the interval $d(z_1,...,z_m)$ corresponding to this hypercube, i.e.,
+\[
+y_i = y(z_1,...,z_m), \; 0 \leq i < 2^{mN}-1.
+\]
+The centers $y_i$ and $y_{i+1}$ correspond to the adjacent hypercubes having a common facet. More details on the indexing of the hypercubes can be found in \cite{6_Strongin2013}.
+\end{enumerate}
+
+Let us consider the mapping  $l(x)$ of the interval $[0,1]$ into a hypercube $D$ defined by the expression 
+\[
+l(x)=y_i+(y_{i+1}-y_i)\frac{w(x)-x_i}{x_{i+1}-x_i}, \; x_i \leq w(x) \leq x_{i+1},
+\]
+\[
+w(x)=x(1-2^{-mN}), \; 0 \leq x \leq 1.
+\]
+The image of any subinterval
+\[
+\left[x_i(1-2^{-mN})^{-1}, x_{i+1}(1-2^{-mN})^{-1}\right], \; 0 \leq i < 2^{mN}-1
+\]
+of the interval $[0,1]$ for the correspondence $l(x)$ is a linear segment connecting the nodes $y_i$ and $y_{i+1}$.
+
+Thus, we have built a piecewise linear curve $l_m(x), 0 \leq x \leq 1,$ connecting the nodes $y_i, 0 \leq i< 2^{mN}-1,$ in the order of their indexing.
+This curve obtained numerically (\textit{the evolvent})  is an approximation of the theoretical Peano curve with the precision not worse than $2^{-m}$ with the respect to each coordinate (the parameter m is called\textit{ the density of the evolvent}).
+As an illustration, an image of the interval $[0,1]$ for the mapping $l_m(x)$ in the case $N=2, m=3$ is presented in Fig.~\ref{6_fig_4}.
+The grid nodes are marked with the dark circles.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.5\linewidth]{figures/6_4.png}
+\caption{A piecewise linear curve}
+\label{6_fig_4}     
+\end{figure}
+
+\section{Multidimensional generalized global search algorithm}
+The considered dimensionality reduction scheme associates a multidimensional problem with Lipschitz objective function and Lipschitz constraints with a one-dimensional problem, where the corresponding functions satisfy uniform H\"older condition (\ref{6_Holder}).
+As result, all lengths of the intervals appearing in the rules of the search algorithm should be replaced by the lengths in a new metrics, where the distance is defined by the expression
+\[
+\Delta_i = \left|x_i-x_{i-1}\right|^{1/N}.
+\]
+Also, in the reduction of the problem dimensionality by Peano curve  $y(x), 0 \leq x \leq 1,$ a trial at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm will consist in the following sequence of operations.
+\begin{itemize}
+	\item Determine the image  $y^k=y(x^k)$ in accordance with the mapping  $y(x)$.
+	\item Compute the values  $g_1(y^k),...,g_\nu(y^k)$, where the index $\nu \leq m$ are determined by the conditions
+	\[
+	g_j(y^k)\leq 0, \; 1 \leq j<\nu, \; g_\nu(y^k)>0, \; \nu \leq m.
+	\]
+	The occurrence of the first violation of the constraint terminates the trial at the point $y^k$.
+In the case, when the point $y^k$ is a feasible one, i.e., when $y(x^k) \in Q_{m+1}$, the trial includes the computation of the values of all functions of the problems and the index is accepted to be $\nu = m+1$.
+The pair of values
+	\begin{equation}\label{6_trial_res}
+	\nu = \nu(x^k), \; z^k = g_\nu(y(x^k))
+	\end{equation}
+	is a \textit{result of the trial} at the point $x^k$.
+\end{itemize}
+
+Thus, one can easily modify the initial one-dimensional algorithm and obtain on this base a multidimensional \textit{generalized index global search algorithm}, described below.
+
+The first trial is executed at an arbitrary internal point $x^1 \in (0,1)$. The selection of the point $x^{k+1}, k \geq 1,$ of any next trial is determined by the following rules.
+
+Rule 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the
+lower indices in ascending order of coordinate values, i.e.
+\begin{equation}\label{6_trial_points}
+0=x_0<x_1<\dots <x_k<x_{k+1}=1,
+\end{equation}
+and juxtapose to them the values $z_i=g_\nu(y(x_i)), \; \nu=\nu(x_i), \; 1 \leq i \leq k$, from (\ref{6_trial_res}) computed at these points. The points $x_0=0$ and
+$x_{k+1}=1$ are introduced additionally for the convenience of further notations, while the values $z_0$ and
+$z_{k+1}$ are not defined.
+
+Rule 2. Classify the indices $i, \; 1 \leq i \leq k$, of the trial points (\ref{6_trial_points})
+according to the number of the problem constraints fulfilled at these points by constructing the sets
+\[
+I_\nu =\left\{i:1 \leq i \leq k, \; \nu=\nu(x_i) \right\}, \; 1 \leq \nu \leq m+1,
+\]
+containing the numbers of all the points $x_i, \; 1 \leq i \leq k$, with
+the same values of $\nu$. The end points $x_0=0$ and $x_{k+1}=1$ are
+interpreted as the ones having indices equal to zero. An additional set $I_0=\left\{0,k+1\right\}$ corresponds to them.
+
+Determine the maximum value of the index
+\[
+M=\max\left\{\nu(x_i), \; 1 \leq i \leq k \right \}.
+\]
+
+Rule 3. Compute the current lower estimates
+\begin{equation}\label{6_mu}
+\mu = \max\left\{ \frac{\left|z_i-z_j\right|}{ (x_i - x_j)^{1/N} }, \; i,j \in I_\nu, \; i>j \right\}
+\end{equation}
+for the unknown H{\"o}lder constants $K_\nu$ of the functions $g_\nu(y),1
+\leq \nu \leq m+1$. If a set $I_\nu$ contains less than two elements, or
+if $\mu_\nu$ from (\ref{6_mu}) is equal to zero, then assume $\mu_\nu=1$.
+
+Rule 4. For all nonempty sets $I_\nu, \; 1 \leq \nu \leq m+1$, compute the
+estimates
+\[
+z_\nu^\ast = \left\{
+   \begin{array}{lr}
+     -\epsilon_\nu, & \nu < M,\\
+     \min\{ g_\nu(y(x_i)): i\in I_\nu \}, & \nu = M,
+   \end{array}
+\right.
+ \]
+where the vector $\epsilon_R = (\epsilon_1,...,\epsilon_m)$ is a predefined vector with the nonnegative components (\textit{the reserve vector}).
+
+Rule 5. For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute
+the \textit{characteristics} $R(i)$ :
+\[
+R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
+\]
+\begin{equation}%\label{eq:14}
+R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
+\end{equation}
+\[
+R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
+\]
+where $\Delta_i=(x_i - x_{i-1})^{1/N}$. The values $r_\nu > 1, \; 1 \leq
+\nu \leq m+1,$ are parameters of the algorithm. An appropriate selection
+of $r_\nu$ allows to consider the product $r_\nu \mu_\nu$ as an estimate
+of the H{\"o}lder constants $K_\nu, \; 1 \leq \nu \leq m+1$.
+
+Rule 6. Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
+\begin{equation}\label{6_MaxR}
+R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
+\end{equation}
+
+Rule 7. Make the next trial at the midpoint of the interval
+$(x_{t-1},x_t)$ if the indices of the points $x_{t-1}$ and $x_t$  are not
+the same, i.e.,
+\[
+x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
+\]
+Otherwise, make the trial at the point
+\begin{equation}%\label{eq:142}
+x^{k+1} = \frac{x_t+x_{t-1}}{2} - \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
+\end{equation}
+
+We may take as termination condition the inequality $\Delta_t \leq
+\epsilon$, where $t$ is from (\ref{6_MaxR}) and $\epsilon>0$ is the
+predefined accuracy.
+
+\textbf{The algorithm convergence conditions.} These conditions are a special case of the convergence theorem which will be proved in Subsection 6.3 for the parallel algorithm with a single curve on a single processor.
+
+\begin{example} \label{6_example1}
+As an illustration, let us consider the problem of minimization of the function
+\begin{align*}
+ \varphi(y_1,y_2) = &-1.5y_1^2\exp(1-y_1^2-20.25(y_1-y_2)^2)- \\
+				            &-(0.5(y_1-1)(y_2-1))^4\exp(2-(0.5(y_1-1))^4-(y_2-1)^4)
+\end{align*}
+within the domain $0 \leq y_1\leq 4,\ -1\leq y_2\leq 3$, with the constraints 
+\begin{align*}
+g_1(y_1,y_2)&=0.01((y_1-2.2)^2+(y_2-1.2)^2-2.25)\leq 0,\\
+g_2(y_1,y_2)&=100(1-(y_1-2)^2/1.44-(0.5y_2)^2)\leq 0,\\
+g_3(y_1,y_2)&=10(y_2-1.5-1.5\sin(6.283(y_1-1.75)))\leq 0.
+\end{align*}
+
+Fig.~\ref{6_fig_5} shows the square search domain and the points of $1098$ trials obtained by the index method with the following parameters: the reliability parameters $r_1=...=r_4=2$, the accuracy $\epsilon = 10^{-3}$, the evolvent density $m=12$.
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.8\linewidth]{figures/6_5.png}
+\caption{The results of solving the test problem}
+\label{6_fig_5}     
+\end{figure}
+
+\end{example}
+
+Note that in applied problems, estimating the values of the constraints and of the objective functions very often requires considerable computation resources.
+The results of the experiments confirm the efficiency of the index scheme of the accounting for the constraints since the values of the functions $g_1,\ g_2,\ g_3,\ \varphi = g_4$ were computed $k_1=1098,\ k_2=623,\ k_3=392,$ and $k_4=152$ times, respectively.
+If the search of the solution were conducted on a uniform grid, in order to achieve the same accuracy $\epsilon=10^{-3}$ $k_1=1.6\cdot 10^7,\ k_2=7 \cdot 10^6,\ k_3=3\cdot 10^6,$ and $k_4=1.4 \cdot 10^6$ computations of the values of the functions $g_1,\ g_2,\ g_3,\ \varphi = g_4$, respectively, would be necessary.
+
+\section{Parallel multidimensional multiextremal methods based on multiple Peano curves}
+
+\subsection{Application of multiple mappings}\label{6_section_shift}
+
+The reduction of the multidimensional problems to the one-dimensional ones using the Peano curves has such important properties as the continuity and the uniform boundedness of the function differences for limited variation of argument.
+However, a partial loss of information on the nearness of the points in the multidimensional space takes place since a point $x \in [0,1]$ has the left and the right neighbors only while the corresponding point $y(x) \in D \subset R^N$ has the neighbors in $2N$ directions.
+As a result, when using the mappings like Peano curve the images $y',\ y'',$ which are close to each other in the $N$-dimensional space can correspond to the preimages $x',\ x'',$ which can be far away from each other in the interval $[0,1]$.
+This property results in the excess computations since several limit points $x',\ x''$ of the trial sequence generated by the index method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional space.
+
+One of the possible ways to overcome this disadvantage consists in using the multiple mappings
+\begin{equation}%\label{eq:142}
+Y_L(x)=\left\{y^0(x),\ y^1(x),...,\ y^L(x)\right\}
+\end{equation}
+instead of single Peano curve $y(x)$ (see \cite{6_Strongin1991,6_Strongin1992,6_Strongin2000}).
+
+Let us consider a family of the hypercubes 
+\begin{equation}\label{6_hypercubes}
+D_l= \left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq L,
+\end{equation}
+where the hypercube $D_{l+1}$ is obtained by the translation of the hypercube $D_l$ along the main diagonal with displacement equal to $2^{-l}$ in each coordinate.
+In Fig.~\ref{6_fig_6} the hypercubes $D_0,...,D_3$ for the case $N=2,\ L=3$ are presented.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.7\linewidth]{figures/6_6.png}
+\caption{Multiple mappings}
+\label{6_fig_6}     
+\end{figure}
+
+Let us assume that the evolvent $y^0(x)$  maps the interval $[0,1]$ onto the hypercube $D_0$ from (\ref{6_hypercubes}), i.e.,
+\[
+D_0 = \left\{y^0(x) : x \in [0,1]\right\}.
+\]
+Then, the evolvents $y^l(x)=\left\{y_1^l(x),...,y_N^l(x)\right\}$, the coordinates of which are defined by the conditions 
+\[
+y_i^l(x)=y_i^{l-1}(x)+2^{-l},\ 1\leq i\leq N, \ 1\leq l\leq L,
+\]
+map the interval $[0,1]$ onto the corresponding hypercubes $D_l,\ 1\leq l \leq L$.
+In Fig.~\ref{6_fig_6} the image of the interval $[0,1]$ obtained by the curve $y^0(x),\ x\in [0,1],$ is shown as the broken line.
+Since the hypercube $D$ from (\ref{6_D}) is included in the common part of the family of hypercubes (\ref{6_hypercubes}) (the boundaries of hypercube $D$ are highlighted in Fig.~\ref{6_fig_6}), having introduced an additional constraint function
+\begin{equation}\label{6_g0}
+g_0(y)=\max\left\{\left|y_i\right| - 2^{-1}:\ 1\leq i\leq N\right\},
+\end{equation}
+one can present the initial hypercube $D$ in the form
+\[
+D=\left\{y^l(x):\; x\in [0,1],\ g_0(y^l(x))\leq 0 \right\},\ 0\leq l \leq L,
+\]
+i.e., $g_0(y) \leq 0$ if $y\in D$ and $g_0(y)>0$ otherwise. Consequently, any point $y \in D$ has its own preimage $x^l \in [0,1]$ for each mapping $y^l(x),\ 0\leq l\leq L$.
+
+Thus, each evolvent $y^l(x),\ 0\leq l \leq L,$ generates its own problem of the type (\ref{6_problem_1}) featured by its own extended (in comparison with $D$) search domain $D_l$ and the additional constraint with the left hand part from (\ref{6_g0})
+\begin{equation}\label{6_problem_l} 
+\min{\left\{\varphi(y^l(x)):x\in [0,1], \; g_j(y^l(x))\leq 0, \; 0 \leq j \leq m\right\}}, \ 0 \leq l \leq L.
+\end{equation}
+
+The problems (\ref{6_problem_l}) correspond to the domains $Q_0^l=[0,1]$ and $Q_{j+1}^l, 0 \leq j \leq m,$ defined by the expression 
+\[
+Q_{j+1}^l = \left\{x \in Q_j^l:g_j(y^l(x))\leq 0\right\},\ 0\leq j\leq m,
+\]
+for the corresponding evolvents $y^l(x)$. The application of a multiple mapping defines the following relation for the nearness in the multidimensional search domain and in the one-dimensional one.
+
+\begin{theorem}
+Let a point $y^\ast$ from the domain $D$ be contained in the line segment with the end-points $y',y'' \in D$ meeting the requirements 
+\[
+\left|y'_j - y''_j\right|\leq 2^{-p},\ y'_i=y''_i=y^\ast_i, \ 1\leq i \leq N,\ i \neq j,
+\]
+where $p$ is an integer  and $1\leq p \leq L$, i.e., the line segment is collinear with the $j$-th axis in $R^N$.
+Then, there exist at least one mapping $y^l(x),\ 0\leq l\leq L$, and the preimages $x^\ast,\; x',\; x''\in [0,1]$ such that
+\[
+y^\ast = y^l(x^\ast),\ y'=y^l(x'),\ y'' = y^l(x'')
+\]
+and
+\[
+\max \left\{ \left|x'-x^\ast\right|,\; \left|x''-x^\ast\right|,\; \left|x'-x''\right|\right\} \leq 2^{-pN}.
+\]
+\end{theorem}
+\begin{proof}
+Proof of this theorem is given in \cite{6_Strongin2000}.
+\qed
+\end{proof}
+
+The conditions of the theorem single out a specific vicinity of the point $y^\ast$.
+This vicinity includes only the points, which can be obtained by the shift of $y^\ast$ parallel to one of the coordinate axes with a displacement not more than $2^{-p}$.
+By changing  $j,\ 1\leq j\leq N,$ in the theorem conditions it is possible to obtain the neighbours in any $N$ coordinate directions.
+According to the statement, the closeness of the points in the $N$-dimensional space in a particular direction will be reflected by the closeness of their preimages in one of the univariate problems.
+The information on the closeness of the points results, first, in more precise estimate of Lipschitz constants and, second, in the increase of the characteristics of the intervals, the images of the end points of which are close to each other in the $N$-dimensional space.
+
+\subsection{Organization of the parallel computations}
+
+Using the multiple mappings allows solving initial problem (\ref{6_problem}) by parallel solving $L+1$ problems of the type (\ref{6_problem_l}) on a set of intervals $[0,1]$ by the index method.
+Each univariate problem is solved on a separate processor.
+The results of trial at the point $x^k$ obtained on a particular processor for the problem being solved by this processor are interpreted as the results of the trials in the rest problems (in the corresponding points $x^{k0},\ x^{k1},...,x^{kL}$).
+In this approach, a trial at the point $x^k \in [0,1]$ executed in the framework of the $s$-th problem, consists in the following sequence of operations.
+\begin{enumerate}
+	\item Determine the image $y^k=y^s(x^k)$ at the mapping  $y^s(x)$.
+	\item Compute the value $g_0(y^k)$.
+If $g_0(y^k) \leq 0$, i.e., if $y^k \in D$, then inform the rest processors on the start of the trial execution at the point $y^k$ (\textit{the blocking} of the point $y^k$).
+	\item Compute the values $g_1(y^k),...,g_\nu(y^k),$ where the index $\nu \leq m$ are determined by the conditions
+	\[
+	g_j(y^k)\leq 0,\ 1\leq j< \nu,\ g_\nu(y^k) > 0,\ \nu \leq m.
+	\]
+	The occurrence of the first violation of any constraint terminates the trial at the point $y^k$.
+In the case when $y^k$ is a feasible one, i.e., when $y^s(x^k) \in Q_{m+1}$, the trial includes the computation of all functions of the problem.
+In this situation the index is set to $\nu = m+1$.
+The triplet
+	\begin{equation}\label{6_triplet} 
+	y^s(x^k),\ \nu = \nu(x^k),\ z^k=g_\nu(y^s(x^k))
+	\end{equation}
+	is  \textit{the trial result} at the point $x^k$.
+	\item If $\nu (x^k)>0$, i.e., if $y^k \in D$, then determine the preimages $x^{kl} \in [0,1],\; 0\leq l \leq L,$ of the point $y^k$ and interpret the trial executed at the point $y^k \in D$ as the execution of the trials in the $L+1$ points
+	\[
+	x^{k0},\ x^{k1},...,\ x^{kL}
+	\]
+	with the same result 
+	\[
+	\nu(x^{k0})=\nu(x^{k1})=...=\nu(x^{kL})=\nu(x^k),	
+	\]
+	\[
+	g_\nu(y^0(x^{k0}))=g_\nu(y^1(x^{k1}))=...=g_\nu(y^L(x^{kL}))=z^k	
+	\]
+	and inform the rest processors on the results of the trial at the point $y^k$.	
+	In the case if $\nu (x^k)=0$, i.e., if $y^k \notin D$, the trial result is related only to the $s$-th problem.	
+\end{enumerate}
+
+Each processor has its own copy of the software realizing the computations of the problem functions and the decision rule of the algorithm.
+For the organization of the interactions among the processors, $L+1$ queues are created on each processor, where the processors store the information on the iterations executed in the form of the triples (\ref{6_triplet}).
+Moreover, the index of the blocked point is assumed to be equal to $–1$; the function value at this point is undefined.
+The links among the processors by means of the queues $Q_{ls}$, where $l,\; 0\leq l \leq L,$ is the number of the transmitting processor and $s,\; 0 \leq s \leq L,$ is the index of the receiving processor, are illustrated in Fig.~\ref{6_fig_7}.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.8\linewidth]{figures/6_7.png}
+\caption{The links among the processors }
+\label{6_fig_7}     
+\end{figure}
+
+The proposed scheme does not include any managing processor that increases the reliability of the computations executed.
+The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the sequential algorithm (except the method of the trial execution).
+However, in the purposes of deeper understanding of the subject, the scheme is presented here in full.
+
+The algorithm for the selection of the iteration points is the same for all processors.
+The starting iteration is executed at a predefined point $x^1 \in (0,1)$ (the starting points for all processors are different).
+The selection of any next iteration point $x^{q+1}$, $q \geq 1$, is determined by the following rules.
+
+Rule 1.
+Extract from all queues associated with $l$-th processor the results stored for this processor including the set of the iteration points $Y_q=\left\{y^{qi}:1 \leq i \leq s_q\right\}$,  the indices and the function values from (\ref{6_triplet}) computed at these points; determine the set $X_q={x^qi:1\leq i \leq s_q}$ of preimages of the points from the set $Y_q$ for the evolvent $y^l(x)$.
+
+Rule 2. Renumber the points of the set of iterations
+\[
+\{x_1\} \cup X_1 \cup ... \cup X_q
+\]
+by the subscripts in increasing order of the coordinate
+\begin{equation}\label{6_points_x} 
+	0=x_0<x_1< \dots <x_k<x_{k+1}=1,
+\end{equation}
+where $k=1+s_1+\dots+s_q$, and juxtapose to them the values $z_i=g_\nu(x_i)$, $\nu = \nu(x_i)$, $1 \leq i \leq k$, computed at these points.
+The index of the blocked point $x_i$ (i.e., of the point, where other trial has been already started by other processor) is set to $–1$, i.e., $\nu(x_i)=-1$ and the value $z_i$ is undefined.
+The points $x_0,\ x_{k+1}$ are introduced additionally for the convenience of further presentation, the indices of these points are set to $-2$, i.e., $\nu(x_0)=\nu(x_{k+1})=-2$ and the values $z_0,\ z_{k+1}$ are undefined.
+
+Rule 3.
+Perform the classification of the subscripts $i,\ 1 \leq i \leq k,$ of the points from (\ref{6_points_x}) in accordance with  the number of problem constraints fulfilled in these points by constructing the sets
+\[
+I_{-2} = \{0, k+1\},
+\]
+\[
+I_{-1} = \{i: 1 \leq i \leq k,\ \nu(x_i)=-1\},
+\]
+\[
+I_{\nu} = \{i: 1 \leq i \leq k,\ \nu(x_i)=\nu\},\ 0 \leq \nu \leq m+1,
+\]
+including the indices of all points $x_i,\ 1 \leq i \leq k,$ with the same indices equal to $\nu$. 
+Determine the maximum value of the index
+\[
+M=\max\left\{\nu(x_i), \; 1 \leq i \leq k \right \}.
+\]
+
+Rule 3. Compute the current lower estimates
+\begin{equation}\label{6_mu_par}
+\mu = \max\left\{ \frac{\left|z_i-z_j\right|}{ (x_i - x_j)^{1/N} }, \; i,j \in I_\nu, \; i>j \right\}
+\end{equation}
+for the unknown H{\"o}lder constants $K_\nu$ of the functions $g_\nu(y),1
+\leq \nu \leq m+1$. If a set $I_\nu$ contains less than two elements, or
+if $\mu_\nu$ from (\ref{6_mu_par}) is equal to zero, then assume $\mu_\nu=1$.
+
+Rule 4. For all nonempty sets $I_\nu, \; 1 \leq \nu \leq m+1$, compute the
+values
+\begin{equation}\label{6_z_nu}
+z_\nu^\ast = \left\{
+   \begin{array}{lr}
+     -\epsilon_\nu, & \nu < M,\\
+     \min\{ g_\nu(y(x_i)): i\in I_\nu \}, & \nu = M,
+   \end{array}
+\right.
+\end{equation}
+where the vector $\epsilon_R = (\epsilon_1,...,\epsilon_m)$ is a predefined vector with the nonnegative components (\textit{the reserve vector}).
+
+Rule 5. For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute
+the \textit{characteristics} $R(i)$ :
+\[
+R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
+\]
+\begin{equation}\label{6_R_par}
+R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
+\end{equation}
+\[
+R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
+\]
+\[
+\Delta_i=(x_i - x_{i-1})^{1/N}
+\]
+The values $r_\nu > 1, \; 1 \leq \nu \leq m+1,$ are parameters of the algorithm.
+
+Rule 6. Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
+\begin{equation}\label{6_MaxR_par}
+R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
+\end{equation}
+
+Rule 7. Make the next trial at the midpoint of the interval
+$(x_{t-1},x_t)$ if the indices of the points $x_{t-1}$ and $x_t$  are not
+the same, i.e.,
+\begin{equation}\label{6_x_new_par_1}
+x^{q+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
+\end{equation}
+Otherwise, make the trial at the point
+\begin{equation}\label{6_x_new_par_2}
+x^{q+1} = \frac{x_t+x_{t-1}}{2} - \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
+\end{equation}
+If $\nu(x^{q+1})=0$, i.e.
+$y^{q+1} \notin D$, then store the trial results only in queue assigned to the current processor itself.
+If $\nu(x^{q+1})>0$, i.e.
+$y^{q+1} \in D$, then store the trial results in all queues assigned to the current processor.
+
+The stopping condition will terminate the search when the inequality
+\[
+\Delta_t \leq \epsilon
+\]
+holds, where $t$ is from (\ref{6_MaxR_par}) and $\epsilon > 0$ is a given search accuracy.
+
+\subsection{The convergence conditions for the algorithm}
+
+The sufficient convergence conditions for the considered parallel algorithm can be formulated in the following form.
+\begin{theorem}\label{6_theorem}
+Assume that the following conditions are satisfied.
+
+\begin{enumerate}
+	\item Problem (\ref{6_problem}) has an $\epsilon$-reserved solution $y_\epsilon$ from (\ref{6_eps_r}).
+	\item Functions $g_j(y),\; 1\leq j\leq m+1,$ admit Lipschitzian (with the constants $L_j$) extensions $G_j(y)$ over the whole feasible domain $D$ from (\ref{6_D}), i.e.,
+	\begin{equation}\label{6_lip_ext}
+	g_j(y(x))=G_j(y(x)),\ x\in Q_j,\ 1\leq j\leq m+1,
+	\end{equation}
+	where the sets $Q_j$ are from (\ref{6_Q}) and $y(x)$ is the curve from (\ref{6_problem_1}).	
+	\item The parameters of the method $\epsilon_\nu,\ 1\leq \nu \leq m,$ used in the rule (\ref{6_z_nu}) are the corresponding components of the reserve vector $\epsilon_R$ from (\ref{6_eps_r}).
+	\item For sufficiently large iteration number $q$  the values $\mu_\nu$ from (\ref{6_mu_par}) satisfy the inequalities 
+	\begin{equation}\label{6_inequalities}
+	r_\nu\mu_\nu > 2^{3-1/N}L_\nu \sqrt{N+3},\ 1\leq \nu \leq m+1,	
+	\end{equation}
+	at least for one processor.
+\end{enumerate}
+Then, any limit point $\overline{y}$  of the sequence of the trials ${y^k}$ generated by the index algorithm for problem (\ref{6_problem}) belongs to the set $Y_\epsilon$ form (\ref{6_eps_solution}) and satisfies the conditions
+	\begin{equation}\label{6_conditions}
+	\varphi(\overline{y})=\inf \left\{\varphi(y^k):g_j(y^k)\leq 0, \; 1\leq j\leq m,\ k=1,2,\dots \right\}\leq \varphi(y_\epsilon).	
+	\end{equation}
+\end{theorem}
+Prior to begin the proof of the theorem let us formulate and prove several auxiliary statements.
+\begin{lemma}
+Assume that the theorem conditions are satisfied. Then, for each limit point $\overline{y}$ of the sequence $\{y^k\}$ there exists an infinite sequence of the intervals 
+	\begin{equation}\label{6_intervals}
+	\{ \left( x_{t-1}, x_t\right): t=t(q^p), p=1,2,...	\},
+	\end{equation}
+	satisfying the conditions 
+	\[
+	\overline{x} = \bigcap_{p=1}^\infty\left[x_{t-1},x_t\right],
+	\]
+	\begin{equation}\label{6_limdelta}
+	\lim_{p\rightarrow\infty}{\Delta_t} = 0,
+	\end{equation}
+	\begin{equation}\label{6_Rg0}
+	R(t(q^p))>0, p=1,2,...,
+	\end{equation}
+	\begin{equation}\label{6_limR}
+	\lim_{p\rightarrow\infty} R(t(q^p))=0,
+	\end{equation}
+where $q^1<q^2<...$; $R(t)$, $\Delta_t$ and $t$ are from the rules (\ref{6_R_par}), (\ref{6_MaxR_par}) correspondingly.
+\end{lemma}
+\begin{proof}
+Since $\overline{y}$ is a limit point, there should exist such index $l,\ 0\leq l \leq L,$ that the $l$-th processor generates a sequence of the points $x^{q^l}$  converging to the preimage $\overline{x}^l$ of the point $\overline{y} = y^l(\overline{x}^l)$ .
+Consequently, there exists a sequence of the trial numbers $q^1,q^2,...,$ satisfying the conditions
+	\begin{equation}\label{6_xl}
+	\overline{x}^l \in \left[x_{t-1},x_t\right],\ t=t(q),\ q\in \{ q^p \},\ p=1,2, ...
+	\end{equation}
+These conditions reflect the fact that the point of the $(q+1)$-th trial executed on the $l$-th processor falls into the interval $[x_{t-1},x_t]$ including the point $\overline{x}^l$ at the step $q$, i.e., $t=t(q)$.
+
+It follows from (\ref{6_mu_par}), (\ref{6_x_new_par_1}), (\ref{6_x_new_par_2}),  and from condition (\ref{6_xl}) that
+	\begin{equation}\label{6_len}
+	\max\left\{x_t-x^{q^l},x^{q^l}-x_{t-1}\right\}\leq \gamma (x_t-x_{t-1}),
+	\end{equation}
+where $\gamma = (r+1)/2r < 1$, $r = \min \{ r_\nu : 1\leq\nu\leq m+1 \} >1$. Then, property (\ref{6_limdelta}) is a consequence of inequality (\ref{6_len}).
+
+At every step $k \geq 1$ there exists an interval $(x_{i-1},x_i): i=i(k)$, at one of the ends of which  the value $z_M^\ast$ is reached, where $M$ is the largest index value.
+If $\nu (x_{i-1}) \neq \nu(x_i)$, according to rules (\ref{6_R_par}), $R(i)=2\Delta_i > 0$.
+In the case, when $\nu(x_{i-1})=\nu(x_i)=M$ the estimate
+\[
+z_i+z_{i-1}-2z_M^\ast \leq \mu_M\Delta_i
+\]
+follows from (\ref{6_mu_par}).
+This estimate in combination with (\ref{6_R_par}) gives the inequality $R(i)=(1-r^{-1})\Delta_i>0$.
+Thus, at any search step $k\geq 1$ an interval with a strictly positive characteristic exists.
+Hence, taking into account (\ref{6_MaxR_par}) one can conclude that the intervals from system (\ref{6_intervals}) should have positive characteristic as well since they undergo partitioning by the trial points falling into these intervals, i.e., conditions (\ref{6_Rg0}) hold.
+
+It follows from the selection rules for the values $z_\nu^\ast$ that
+\[
+z_j = g_\nu(x_j)\geq z_\nu^\ast,\ \nu=\nu(x_j),\ 1\leq j\leq k.
+\]
+Then, from the rules for the computation of the characteristics (\ref{6_R_par}) taking into account conditions (\ref{6_lip_ext}) and inequality (\ref{6_Rg0}) one can derive out the truth of (\ref{6_limR}).
+\end{proof}
+
+\begin{lemma} \label{6_lemma2}
+(on the existence of a feasible point). 
+Assume the conditions of theorem are satisfied. Then, there exists such number of trial $h \geq 1$, for which 
+	\begin{equation}\label{6_fis_point}
+	\nu(x^h)=m+1.
+	\end{equation}
+\end{lemma}
+\begin{proof}
+According to definition (\ref{6_eps_r}), the point $y_\epsilon$ is an internal point of a feasible domain.
+Consequently, for any mapping $y^l(x),\; 0\leq l \leq L,$ there exists a hypercube $D(m)$ of the $m$-th partitioning belonging to a feasible domain and including the $\epsilon$-reserved solution $y_\epsilon$.
+It means that there exists also a preimage interval $d(m)$ for the hypercube $D(m)$ and this interval includes the preimage $x_\epsilon$ of the point $y_\epsilon$, i.e., $y_\epsilon=y^l(x_\epsilon)$.
+At any search step $q>1$ there exists an interval $[x_{j-1},x_j],\; j=j(q),$ including the point $x_\epsilon$.
+
+It follows from H\"older properties (\ref{6_Holder}) of the functions and from condition (\ref{6_lip_ext}) of the theorem that
+	\begin{equation}\label{6_39}
+	z_j=g_\nu(y^l(x_j))\leq g_\nu(y^l(x_\epsilon))+2L_\nu \sqrt{N+3}(x_j - x_\epsilon)^{1/N},\ \nu=\nu(x_j),
+	\end{equation}
+	\begin{equation}\label{6_40}
+	z_{j-1}=g_\nu(y^l(x_{j-1}))\leq g_\nu(y^l(x_\epsilon))+2L_\nu \sqrt{N+3}(x_\epsilon -x_{j-1})^{1/N},\ \nu=\nu(x_{j-1}).
+	\end{equation}
+Moreover, it follows from condition (\ref{6_eps_r}) that
+	\begin{equation}\label{6_41}
+	g_\nu(y(x_\epsilon))\leq -\epsilon_\nu,\ 1\leq \nu \leq m.
+	\end{equation}
+If condition (\ref{6_fis_point}) is not satisfied at any of the search steps, then
+\[
+\nu = \max \{ \nu(x_{j-1}),\nu(x_j) \} \leq m,
+\]
+and it follows from conditions (\ref{6_R_par}), (\ref{6_inequalities}), and (\ref{6_39})--(\ref{6_41}) that at large enough $q$
+\[
+R(j(q)) \geq \Delta_j \left(1-\frac{4L_\nu\sqrt{N+3}}{r_\nu\mu_\nu}\right) + 4\frac{z_\nu^\ast+\epsilon_\nu}{r_\nu\mu_\nu}>0, \ \nu(x_{j-1})=\nu(x_j),
+\]
+\[
+R(j(q)) \geq 2\Delta_j \left(1-\frac{2L_\nu\sqrt{N+3}}{r_\nu\mu_\nu}\right) + 4\frac{z_\nu^\ast+\epsilon_\nu}{r_\nu\mu_\nu}>0, \ \nu(x_{j-1}) \neq \nu(x_j).
+\]
+
+From these inequalities, from limit (\ref{6_limR}) as well as from rule (\ref{6_MaxR_par}) a necessity to execute the trials in the considered interval $[x_{j-1},x_j]$ at some search step follows, that results in its splitting into two subintervals.
+At least one of these includes the point $x_\epsilon$.
+The repetition of this process for $q\rightarrow \infty$ generates an infinite sequence of the subintervals collapses to the point $x_\epsilon$.
+It has been shown in the beginning of the proof of this lemma that there exists an interval $d(m)$ including the point $x_\epsilon$ and that all points of this interval are the internal points of a feasible domain, i.e., their indices are equal to $m+1$.
+Consequently, the interval $[x_{j-1},x_j]$ from the sequence of those collapsing to $x_\epsilon$ will belong to the interval $d(m)$.
+This proves the truth of (\ref{6_fis_point}).
+\end{proof}
+
+\begin{lemma} (on the feasibility of any limit point).
+Any limit point $\overline{y}=y^l(\overline{x}^l)$  of the trial sequence $\{y^k\}$ is feasible, i.e., 
+	\begin{equation}\label{6_rho}
+	\nu(\overline{x}^l)=m+1.
+	\end{equation}
+\end{lemma} 
+\begin{proof}
+Let us assume that the limit point $\overline{y}$ is not feasible, i.e. that there exists such integer $\nu \geq 1$ that
+\[
+g_i(y(\overline{x}^l))\leq 0,\ 1\leq i\leq \nu - 1,\ g_\nu(y(\overline{x}^l))>0.
+\]
+Then, the point $\overline{y}$ has a vicinity $U_\rho(\overline{y})$ of the radius $\rho = \epsilon_\nu/2L_\nu$ such that all points of this vicinity satisfy the condition
+	\begin{equation}\label{6_lim_fis_point}
+	g_\nu(y) > -\epsilon_\nu/2, \ y \in U_\rho(\overline{y}).
+	\end{equation}
+
+If an integer number $M$ satisfies the inequality $2^{-M\sqrt{N}}<\rho$, there exists a subcube $D(M)$ including the point  $\overline{y}$.
+If $\overline{y}$ is an internal point of $D(M)$ then the preimage of this subcube $d(M)$ includes all preimages of the point $\overline{y}$.
+Consequently, $\overline{x}$ is an internal point of $d(M)$.
+If $\overline{y}$ is a corner of the subcube $D(M)$, there exists $2^N$ subcubes of the $M$-th  partitioning in $U_\rho(\overline{y})$, and their preimages include all preimages of the point $\overline{y}$ .
+It means that $\overline{x}^l$  belongs to one of the preimages.
+Let us denote this interval as  $d'(M)$.
+If $\overline{x}^l$ is an end point of $d'(M)$, this point is an end point of another interval $d''(M)$ as well.
+Consequently, the union $d'(M) \cup d''(M)$ includes $\overline{x}^l$ as an internal point. 
+Thus, in any case there exists an interval $[\alpha,\beta] \subset [0,1] $ including $\overline{x}^l$ as an internal point and the images $y=y(x)$ of all points of the interval $(\alpha,\beta)$ satisfy condition (\ref{6_lim_fis_point}).
+
+Since  $\overline{x}^l$  is a limit point, the intervals from the sequence $[x_{t-1},x_t],\ t=t(q)$, collapsing to this point should satisfy the condition $[x_{t-1},x_t] \subset (\alpha,\beta)$ for sufficiently large $q$.
+Thus, it follows from rule (\ref{6_z_nu}) and statement (\ref{6_fis_point}) that at any $q \geq h$
+	\begin{equation}\label{6_44}
+	z_\nu^\ast = -\epsilon_\nu,\ 1\leq\nu\leq m,
+	\end{equation}
+and from (\ref{6_R_par}), (\ref{6_lim_fis_point}), and (\ref{6_44}) we obtain the estimate 
+\[
+R(t(q)) \leq \Delta_i + \frac{(z_i - z_{i-1})^2}{r_\nu^2\mu_\nu^2\Delta_i}-\frac{2\epsilon_\nu}{r_\nu\mu_\nu}.
+\]
+Consequently, according to (\ref{6_limdelta}) for sufficiently large $q$
+	\begin{equation}\label{6_45}
+R(t(q)) < \frac{-\epsilon_\nu}{r_\nu\mu_\nu} < 0.
+	\end{equation}
+The last inequality contradicts to condition (\ref{6_Rg0}) that in turn leads to a contradiction with the limiting character of the point $\overline{x}$.
+\end{proof}
+\begin{proof} (of the theorem \ref{6_theorem})
+
+In order to justify the left equality in (\ref{6_conditions}) let us prove the truth of the inequalities 
+	\begin{equation}\label{6_46}
+	\varphi(y(\overline{x})) \leq z_{m+1}^\ast=\min\{z_i:i\in I_{m+1} \}
+	\end{equation}
+for a feasible limit point $\overline{x}$ of the sequence $\{x^k\}$. Assume that the sequence of the nested intervals (\ref{6_intervals}) at any $p \geq 1$ does not satisfy the conditions
+	\begin{equation}\label{6_47}
+	\max \{\nu(x_{t-1},\nu(x_t)\}=m+1,\ t=t(q^p).
+	\end{equation}
+Then, the truth of inequalities (\ref{6_45}) follows from the inequalities 
+\[
+g_\nu(y(x_i))>0,\ \nu = \nu(x_i),\ i=t-1,t,
+\]
+the rules (\ref{6_R_par}) and the relation (\ref{6_44}) for large enough $q$, that contradicts to the condition of positiveness of characteristic (\ref{6_Rg0}).
+This means that condition (\ref{6_47}) is satisfied (for sufficiently large iteration number $q$).
+
+If (\ref{6_46}) does not take place, for some positive number $\delta$ the inequalities 
+\[
+z_i=g_{m+1}(y(x_i)) > z_{m+1}^\ast + \delta,\ \nu(x_i)=m+1,\ i=t,t-1,\ t=t(q^p),
+\]
+should be fulfilled. From these inequalities, rules for the characteristic (\ref{6_R_par}), condition (\ref{6_47}), and limit (\ref{6_limR}) the estimate for the characteristic 
+\[
+R(t(q^p))\leq 2\Delta_t - \frac{4\delta}{r_{m+1}\mu_{m+1}}<-\frac{4\delta}{r_{m+1}L_{m+1}} < 0 
+\]
+can be obtained, if the number $p$ is large enough. However, this estimate contradicts to the positiveness of the characteristic (\ref{6_Rg0}). Consequently, the property (\ref{6_46}) takes place.
+
+In order to prove the right inequality in (\ref{6_conditions}), let us demonstrate that the relation
+\[
+\varphi (y_\epsilon) \leq z_{m+1}^\ast - \beta,
+\]
+is not true for sufficiently large $q$ if $\beta>0$.
+
+When proving Lemma \ref{6_lemma2} the interval  $[x_{j-1},x_j],\ j=j(q)$, was considered, which includes the preimage $x_\epsilon$ of the point $y_\epsilon$ and for which condition (\ref{6_47}) is satisfied if iteration number $q$ is sufficiently large.
+Then, from the condition
+\[
+\varphi (y_\epsilon) = g_{m+1}(y(x_\epsilon)) \leq z_{m+1}^\ast - \beta,
+\]
+and also from (\ref{6_R_par}), (\ref{6_39}), and (\ref{6_39}) it follows that
+\[
+R(j(q)) \geq \Delta_j \left(1-\frac{4L_{m+1}\sqrt{N+3}}{r_{m+1}\mu_{m+1}}\right) + \frac{4\beta}{r_{m+1}\mu_{m+1}}>0, \ \nu(x_{j-1})=\nu(x_j),
+\]
+\[
+R(j(q)) \geq 2\Delta_j \left(1-\frac{2L_{m+1}\sqrt{N+3}}{r_{m+1}\mu_{m+1}}\right) + \frac{4\beta}{r_{m+1}\mu_{m+1}}>0, \ \nu(x_{j-1}) \neq \nu(x_j).
+\]
+From these inequalities, from H\"older condition (\ref{6_Holder}) and from the theorem assumption (\ref{6_inequalities}) the estimate 
+\[
+R(j(q)) > \frac{4\beta}{r_{m+1}\mu_{m+1}} > \frac{\beta}{r_{m+1}L_{m+1}\sqrt{N+3}} > 0 
+\]
+valid for large enough $q$ can be obtained.
+
+On one hand, according to this inequality and rule (\ref{6_MaxR_par}), the interval $[x_{j-1},x_j]$ should shrink to a point.
+On the other hand, the shrinking interval should satisfy condition (\ref{6_limR}) that contradicts to the existence of the limit.
+Consequently, the lower estimate $z_{m+1}^\ast$  coinciding with the middle part of relation  (\ref{6_conditions}) cannot exceed the value of $\varphi(y_\epsilon)$.
+
+Theorem has been proved.
+
+\end{proof}
+
+\begin{example}
+Let us consider the problem from Example \ref{6_example1} and solve it by parallel index algorithm with the following parameters: the reliability parameters $r_1=...=r_4=1.4$, the search accuracy $\epsilon=10^{-3}$, the evolvent density $m=12$, the reserving parameters $\epsilon_\nu=10^{-2}$.
+The number of processors and, correspondingly, the number of evolvents were varied from 2 to 6.
+The same estimate of the optimum $(y_1^\ast,y_2^\ast) =(0.942, 0.945)$ were obtained in all experiments.
+
+Fig.~\ref{6_fig_8} presents the square search area, where the feasible domain of the problem is highlighted and the trial points obtained by the parallel index method using multiple mapping with six evolvents (and, consequently, solved by six processors) are marked with dots.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.8\linewidth]{figures/6_8.png}
+\caption{The results of solving the problem using six processors}
+\label{6_fig_8}     
+\end{figure}
+\end{example}
+
+The results of the experiments (the number $k_i$ of computations of the values of the problem functions $g_i(y)$) are presented in tables \ref{6_tab2}--\ref{6_tab4}.
+Table~\ref{6_tab1} corresponding to the sequential algorithm is presented for comparison.
+
+\begin{table}
+	\caption{The results of solving the problem on single processor}
+	\label{6_tab1}
+	\center
+	\begin{tabular}{ccccc}
+		\hline\noalign{\smallskip}
+		$l$  & $k_1$ & $k_2$ & $k_3$ & $k_4$    \\
+		\noalign{\smallskip} \hline \noalign{\smallskip}
+		1 	&	1098 &	623 & 392	& 152 \\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}
+	\caption{The results of solving the problem on two processors}
+	\label{6_tab2}
+	\center
+	\begin{tabular}{cccccc}
+		\hline\noalign{\smallskip}
+		$l$  & $k_0$ &$k_1$ & $k_2$ & $k_3$ & $k_4$    \\
+		\noalign{\smallskip} \hline \noalign{\smallskip}
+		1 	&	636 &	567 & 331	& 212 & 104 \\
+		2 	&	632 &	327 & 193	& 125 & 61 \\
+		\noalign{\smallskip}\hline\noalign{\smallskip}
+		Total 	&	1268 &	894 & 524	& 337 & 165 \\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}
+	\caption{The results of solving the problem on four processors}
+	\label{6_tab3}
+	\center
+	\begin{tabular}{cccccc}
+		\hline\noalign{\smallskip}
+		$l$  & $k_0$ &$k_1$ & $k_2$ & $k_3$ & $k_4$    \\
+		\noalign{\smallskip} \hline \noalign{\smallskip}
+1	&	585	&	508	&	311	&	206	&	88	\\
+2	&	579	&	262	&	157	&	86	&	36	\\
+3	&	579	&	293	&	174	&	119	&	41	\\
+4	&	579	&	284	&	178	&	116	&	48	\\
+		\noalign{\smallskip}\hline\noalign{\smallskip}
+		Total	&	2322	&	1347	&	820	&	527	&	213 \\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}
+	\caption{The results of solving the problem on six processors}
+	\label{6_tab4}
+	\center
+	\begin{tabular}{cccccc}
+		\hline\noalign{\smallskip}
+		$l$  & $k_0$ &$k_1$ & $k_2$ & $k_3$ & $k_4$    \\
+		\noalign{\smallskip} \hline \noalign{\smallskip}
+1	&	354	&	289	&	165	&	105	&	44	\\
+2	&	346	&	90	&	55	&	28	&	10	\\
+3	&	346	&	113	&	66	&	40	&	10	\\
+4	&	346	&	160	&	87	&	48	&	25	\\
+5	&	346	&	224	&	119	&	68	&	23	\\
+6	&	346	&	274	&	159	&	97	&	37	\\
+		\noalign{\smallskip}\hline\noalign{\smallskip}
+Total	&	2084	&	1150	&	651	&	386	&	149	\\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+The iteration speedup of the parallel algorithm on $p$ processors is presented in Table~\ref{6_tab5}.
+The iteration speedup was defined as the ratio of the number of iteration (which is equal to the number of checks of the first constraint $g_1$) executed by the sequential algorithm to the maximum number of iterations executed by the parallel algorithm on one of processors.
+This method of determining the speedup is important for the problems, in which the estimates of the functions' values requires considerable computational costs.
+
+\begin{table}
+	\caption{Iteration speedup of the parallel algorithm }
+	\label{6_tab5}
+	\center
+	\begin{tabular}{cc}
+		\hline\noalign{\smallskip}
+		$p$  & speedup     \\
+		\noalign{\smallskip} \hline \noalign{\smallskip}
+		1 	&	--- \\
+		2 	&	1.93 \\
+		4 	&	2.16 \\
+		6 	&	3.8 \\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+The results of the experiments clearly demonstrate the effect of speed up (the reduction in the number of the costly computations of the problem functions).
+
+\section{Parallel computations based of novel schemes for building the multiple Peano curves}
+\subsection{Scheme for building the rotated evolvents}
+
+The application of the scheme for building the multiple evolvents (hereinafter called the shifted evolvents or $S$-evolvents) described in Subsection \ref{6_section_shift} allows to preserve the information on the nearness of the points in the multidimensional space and, therefore, to provide more precise (as compared to a single evolvent) estimate of Lipschitz constant in the search process.
+However, this approach has serious restrictions, which narrow the applicability of the parallel algorithms, designed on the base of the $S$-evolvents.
+
+Because a shifted evolvent is built as a result of the shift of a hypercube $D$ along the main diagonal, and the shift step decreases 2 times for the building of each subsequent mapping, the number of such shifts is limited by the evolvent density $m$.
+In the case when the shift step is less than $2^{-m}$, the next evolvent coincides with the previous one.
+Thus, the applicable number of the evolvents $L$ and, hence, the number of processors are limited by $m$ ($L\leq m$), where $m$ is the density of the evolvents.
+
+Also, it follows from the algorithm for building the shifted evolvents that solving the initial problem in the domain $D$ is reduced to solving a set of problems in the extended domains $D_l$ from (\ref{6_hypercubes}) with additional constraint (\ref{6_g0}).
+As a result, the structure of the search domain becomes more complicated, and the exponential decrease of the volume of the search domain $D$ in relation to the volumes of the domains $D_l$ with increasing dimensionality of the problem takes place.
+
+To overcome the disadvantages mentioned above and to preserve the information on the nearness of the points in the $N$-dimensional space, a novel scheme of building of the multiple mappings is proposed.
+The building of a set of Peano curves not by the shift along the main diagonal of the hypercube but by rotation of the evolvents around the coordinate origin is a distinctive feature of the proposed scheme \cite{6_Gergel2009}.
+In the initial non-rotated mapping for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other.
+In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer.
+The evolvents generated according to the novel scheme will be hereinafter called the rotated evolvents or $R$-evolvents.
+
+In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case $N=2$ are presented as an illustration.
+The grid nodes in the $N$-dimensional space are marked by the dark dots.
+A pair of points, the preimages of which are far away from each other on the one-dimensional axis for one mapping and are close to each other for another mapping is pointed by the arrows.
+The maximum number of various rotations of the evolvents mapping the $N$-dimensional hypercube onto a one-dimensional interval is $2^N$.
+The employment of all possible rotations might appear to be redundant.
+In this case, one can select only a part of all rotations.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.8\linewidth]{figures/6_9.png}
+\caption{Close/far points for the rotated evolvents}
+\label{6_fig_9}     
+\end{figure}
+
+As a possible approach, one can propose to generate the new evolvents by the rotation of the initial evolvents on the angle of $\pm\pi/2$ in each of the coordinate planes.
+For example, the rotation matrices in the plane $(y_2,y_4)$ for  $N=5$ are presented below.
+\begin{eqnarray*} 
+ \begin{pmatrix}
+  1 & 0 & 0 & 0 & 0 \\
+  0 & 0 & 0 & -1 & 0 \\
+  0 & 0 & 1 & 0 & 0 \\
+  0 & 1 & 0 & 0 & 0 \\
+  0 & 0 & 0 & 0 & 1
+ \end{pmatrix}
+\qquad \qquad
+ \begin{pmatrix}
+  1 & 0 & 0 & 0 & 0 \\
+  0 & 0 & 0 & 1 & 0 \\
+  0 & 0 & 1 & 0 & 0 \\
+  0 & -1 & 0 & 0 & 0 \\
+  0 & 0 & 0 & 0 & 1
+ \end{pmatrix}
+\end{eqnarray*}
+
+The number of such pairs of rotations is determined by the number of the coordinate planes in the space, which equals to  $N(N-1)/2$.
+The total number of the transformations equals to $N(N-1)$.
+Taking into account the initial mapping, one can conclude that this method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals.
+Moreover, the additional constraint  $g_0(y) \leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent.
+This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.
+
+The use of the set of mappings $Y_L(x)=\{y^1(x),...,y^L(x)\}$ results in the appearing of the corresponding set of the one-dimensional multiextremal problems 
+\begin{equation}\label{6_problem_lr} 
+\min{\left\{\varphi(y^l(x)):x\in [0,1], \; g_j(y^l(x))\leq 0, \; 1 \leq j \leq m\right\}}, \ 1 \leq l \leq L.
+\end{equation} 
+Each problem from this set can be solved independently.
+Any computation of the value $z=g_\nu(y'),\ y'=y^i(x')$ of the function $g_\nu(y)$ in the $i$-th problem can be interpreted as a computation of the value $z=g_\nu(y'),\ y'=y^s(x'')$ for any other $s$-th problem without time-comsuming computations of the functions $g_\nu(y)$.
+Such information integrity allows to solve  initial problem (\ref{6_problem}) by solving $L$ problems (\ref{6_problem_lr}) on a set of intervals $[0,1]$ in parallel by the index method.
+
+The decision rules for the parallel algorithm using $R$-evolvents coincide with the ones of the sequential algorithm almost completely except the method of the trial execution.
+Namely, the execution of a trial at the point $x^k \in [0,1]$ in the $s$-th problem in the case of using $R$-evolvents consists in the following sequence of operations.
+
+\begin{enumerate}
+	\item Determine the image $y^k=y^s(x^k)$ according to the mapping  $y^s(x)$.
+	\item Inform the rest processors on the beginning of the trial execution at the point  $y^k$ (\textit{the blocking} of the point $y^k$).
+	\item Compute the values $g_1(y^k),...,g_\nu(y^k),$ where the index $\nu \leq m$ are determined by the conditions
+	\[
+	g_j(y^k)\leq 0,\ 1\leq j< \nu,\ g_\nu(y^k) > 0,\ \nu \leq m.
+	\]
+	The first violation of the constraints terminates the trial at the point  $y^k$.
+In the case when $y^k$ is a feasible one, the trial includes the computation of all functions of the problem, and the index value is set to $\nu = m+1$.
+The triplet
+	\begin{equation}\label{6_triplet_r} 
+	y^s(x^k),\ \nu = \nu(x^k),\ z^k=g_\nu(y^s(x^k))
+	\end{equation}
+	is  \textit{the trial result} at the point $x^k$.
+	\item Determine the preimages $x^{kl} \in [0,1],\; 1\leq l \leq L,$ of the point $y^k$ and interpret the trial executed at the point $y^k \in D$ as the execution of the trials in the $L$ points 
+	\[
+	x^{k1},...,\ x^{kL}
+	\]
+	with the same result 
+	\[
+	\nu(x^{k1})=...=\nu(x^{kL})=\nu(x^k),	
+	\]
+	\[
+	g_\nu(y^1(x^{k1}))=...=g_\nu(y^L(x^{kL}))=z^k	
+	\]
+	and inform the rest processors on the results of the trial at the point $y^k$.		
+\end{enumerate}
+
+\subsection{Comparison of the algorithms}
+
+\begin{example}
+
+In order to compare the efficiency of the algorithms, the method of the operation characteristics has been applied.
+The series of experiments was devoted to minimixation of 100 two-dimensional functions of two variables
+\begin{eqnarray} \label{6_VAG}
+\varphi(y)= -&\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2+\right. \\
+&\left.\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},\\ \nonumber
+\end{eqnarray}
+where
+\begin{eqnarray} \nonumber
+& y=(y_1,y_2)\in R^2, 0 \leq y_1,y_2 \leq 1, \\ \nonumber
+& g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2),  \\ \nonumber
+& h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2), \nonumber 
+\end{eqnarray}
+and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
+
+The experiments have been carried out for the algorithms described above with $S$-evolvents and $R$-evolvents  with the density $m = 12$.
+The number of evolvents $L = 2$, the parameters $r=2.1,\ \epsilon = 0.01$ and the reserves $\epsilon_\nu= 0.05$ from (\ref{6_z_nu}) have been used.
+These parameters are the minimal ones for the corresponding methods, for which the $100\%$ of the problems was solved.
+
+The operation characteristics of the methods obtained for the problem series generated according to (\ref{6_VAG}) are presented in Fig.~\ref{6_fig_11}.
+The lower dashed curve corresponds to the shifted evolvents method, the middle dashed one corresponds to the method with a single evolvent, and the upper solid curve corresponds to the rotated evolvents method.
+Such position of the curves demonstrates that the algorithm with $R$-evolvents provides in average faster obtaining the trial points falling into the predefined vicinity of the solution than the algorithm with $S$-evolvents  or the algorithm with a single evolvent.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.8\linewidth]{figures/6_11.png}
+\caption{The operation characteristics}
+\label{6_fig_11}     
+\end{figure}
+
+\end{example}
+
+\begin{example}
+To illustrate the work of the parallel algorithm with $R$-evolvents for solving the essentially multidimensional problems, let us consider the minimization problem of the function
+\[
+\varphi(y) = \sum_{i=1}^N{\left(y_i^2-cos(18y_i^2)\right)},\ -1.5\leq y_i\leq 1.5, \ 1\leq i\leq N,\ N=6,
+\]
+where the minimum value $\varphi(y^\ast) = -N$ is achieved at the point $y^*=0$.
+
+To solve this problem, the sequential method and the parallel one with the rotated evolvents with the density $m = 10$ were used.
+The parameter of the method $r=2.0$ and accuracy $\epsilon = 0.05$ in the termination criterion were employed.
+The number of evolvents $L = 30$ corresponds to the number of processors involved in solving the problem.
+The sequential algorithm as well as the parallel one have found the solution with the required precision; the sequential algorithm has executed 173116 iterations while the parallel one has executed 8535 (the maximum number of iterations for one processor).
+The iteration speedup was 20.28 and the time speedup was 7.48.
+\end{example}
+
+\begin{example}
+Let us consider now the problem of minimization of the function from the previous example in the domain  $-1\leq y_i \leq 1,\ 1\leq i\leq N,$ for $N=15, 20$.
+In this experiment 100 processors were used.
+The speedup in time and the efficiency of finding the global optimum with respect to the number of trials with equal termination conditions were estimated.
+The results of solving the problems are presented in Table~\ref{6_tab_l1}.
+
+	\begin{table}
+	\caption{The results of multidimensional functions minimization}
+	\label{6_tab_l1}
+	\center
+	\begin{tabular}{cccccccc}
+		\hline\noalign{\smallskip}
+		$N$ & $k$ & \multicolumn{2}{c}{ Sequential algorithm  } &\ & \multicolumn{3}{c}{Parallel algorithm } \\
+		\noalign{\smallskip} \cline{3-4} \cline{6-8} \noalign{\smallskip}
+		 & & Time & $\varphi_k^\ast$  &\  & Time & $\varphi_k^\ast$ & Speedup  \\
+		\noalign{\smallskip} \hline \noalign{\smallskip}
+%		11	&	$5\cdot10^6$	&	544068.8 & $-10.997$	& &	9321.5 & $-10.995$ 	&	112.01	\\
+		15\	&	$3\cdot10^6$ \ &	605788 & $-14.835$	& &	7414 & $-14.925$ 	&	81.7	\\
+		20\	&	$3\cdot10^6$ \ &	605956 & $-19.989$	& &	6654 & $-19.048$ 	&	91.06	\\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+It is clear from Table~\ref{6_tab_l1} that the index method with $R$-evolvents executed on 100 processors demonstrated the speedup close to the linear one with respect to the computational time that gives evidence its high efficiency.
+\end{example}
+\begin{example}
+Now let us demonstrate the efficiency of the parallel index method in solving the constrained global optimization problem
+\[
+\varphi(y) = N + \sum_{i=1}^N{\left(y_i^2-\cos(2\pi y_i)\right)}\rightarrow\min 
+\]
+\[
+g_1(y) = \sum_{i=1}^N{y_i} - 0.5 \leq 0
+\]
+\[
+g_2(y) = \sum_{i=1}^N{y_i^2} - 1.0 \leq 0
+\]
+\[
+g_3(y) = \sum_{i=1}^N{\sin(i\pi y_i)} - 0.3 \leq 0
+\]
+for $-1\leq y_i\leq 1,\ 1\leq i\leq N,\ N=8$.
+The results of solving this problem by the sequential method and by the parallel method on 56 processors are presented in Table~\ref{6_tab_l2}.
+In both cases, the methods terminated the search process upon achievement the predefined accuracy $\epsilon = 0.0035$.
+Table~\ref{6_tab_l2} shows number of trials $k$, number of objective function evaluations $k_4$, time of comutations $t$ in seconds, speedup in time $s$, iterations $S$ and in objective function evaluations $S_4$.
+The maximum number of trials and objective function evaluations per one processor is specified in brackets.
+
+	\begin{table}
+	\caption{Comparison of the algorithms in solving a constrained problem}
+	\label{6_tab_l2}
+	\center
+	\begin{tabular}{cccccccc}
+		\hline\noalign{\smallskip}
+		 Method & $k$ & $k_4$ & $S$ & $S_4$ & $t$ & $s$ & $\varphi_k^\ast$  \\
+		\noalign{\smallskip} \hline \noalign{\smallskip}
+		Sequential	&	309640 &	33874 & ---	& --- &	40006 & ---	&	0.016	\\
+		Parallel	&	244356 (4891) \ &	22722 (789) & 63.3	& 42.9 &	626 & 63.8 	&	0.022	\\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+It is evident from Table~\ref{6_tab_l2} that the parallel algorithm demonstrated the speed up 63.3 with respect to the number of trials, 42.9 with respect to the number of objective function computations, and of 63.8 with respect to the computation time.
+Note that the comparison of the sequential algorithm with the parallel one appears to be more difficult in the problems of higher dimensionality since in the latter case all the search information is concentrated in one computational node.
+There appeared to be not enough RAM for the sequential algorithm run on a single node.
+As a result, the swapping was activated.
+In the parallel algorithm, all data are distributed among the processors.
+This allows not to use the external memory for storing the search information.
+
+\end{example}
+
+\begin{thebibliography}{99.}
+
+\bibitem{6_Sergeyev2006}
+Sergeyev, Ya.D., Kvasov, D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim. \textbf{16(3)}, 910--937 (2006)
+
+\bibitem{6_Sergeyev2015}
+Sergeyev, Ya.D., Kvasov, D.E.: A deterministic global optimization using smooth diagonal auxiliary functions. Communications in Nonlinear Science and Numerical Simulation. \textbf{21}, 99--111 (2015)
+
+\bibitem{6_Sergeyev2017}
+Sergeyev, Ya.D., Kvasov, D.E.: Deterministic global optimization: An introduction to the diagonal approach. Springer, New York (2017)
+
+\bibitem{6_Zilinskas2008}
+\v{Z}ilinskas, J.: Branch and bound with simplicial partitions for global optimization. Mathematical Modelling and Analysis \textbf{13(1)}, 145--159 (2008)
+
+\bibitem{6_Zilinskas2014}
+Paulavi\v{c}ius, R., \v{Z}ilinskas, J.: Simplicial Lipschitz optimization without the Lipschitz constant. J. Glob. Optim. \textbf{59(1)}, 23--40  (2014)
+
+\bibitem{6_Zilinskas2014_1}
+Paulavi\v{c}ius, R., \v{Z}ilinskas, J.: Simplicial Global Optimization. SpringerBriefs in Optimization, Springer, New York (2014)
+
+\bibitem{6_Strongin2000}
+Strongin, R.G., Sergeyev, Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000)
+
+\bibitem{6_Strongin2013}
+Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting Space-Filling Curves. SpringerBriefs in Optimization, Springer, New York (2013)
+
+\bibitem{6_Gourdin}
+Gourdin, E., Jaumard, B., Ellaia, R.: Global optimization of H\"{o}lder functions. J. Global Optim. \textbf{8}, 323--348 (1996)
+
+\bibitem{6_Lera2002}
+Lera, D., Sergeyev, Y.D.: Global minimization algorithms for H\"{o}lder functions. BIT \textbf{42(1)}, 119--133 (2002)
+
+\bibitem{6_Lera2010}
+Lera, D., Sergeyev, Y.D.: Lipschitz and H\"{o}lder global optimization using space-filling curves. Appl. Numer. Math. 60(1–2), 115–129 (2010)
+
+\bibitem{6_Hime}
+e Oliveira, H.A., Petraglia, A.: Global optimization using space-filling curves and measure-preserving transformations. Soft Computing in Industrial Applications. \textbf{96}, 121--130 (2011)
+
+\bibitem{6_Lera2015}
+Lera, D., Sergeyev, Y.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and Hölder constants.
+Commun.
+Nonlinear Sci.
+Numer.
+Simul.
+\textbf{23}, 328--342 (2015)
+
+\bibitem{6_Butz}
+Butz, A.R.: Space filling curves and mathematical programming. Inform. Control \textbf{12(4)}, 314--330 (1968)
+
+\bibitem{6_Sagan}
+Sagan, H.: Space–Filling Curves. Springer–Verlag, New York (1994)
+
+\bibitem{6_Strongin1991}
+Strongin, R.G.: Parallel multi-extremal optimization using a set of evolvents. Comp. Math. Math. Phys. \textbf{31(8)}, 37--46 (1991)
+
+\bibitem{6_Strongin1992}
+Strongin, R.G.: Algorithms for multi-extremal mathematical programming problems employing the set of joint space-filling curves. J. Global Optim. \textbf{2(4)}, 357--378 (1992)
+
+\bibitem{6_Gergel2009}
+Strongin, R.G., Gergel, V.P., Barkalov, K.A.: Parallel methods for global optimization problem solving. Journal of instrument engineering. \textbf{52}, 25--33 (2009) (In Russian)
+
+\end{thebibliography}
 

### 2017-10-04 2017/ParOptBook/Chapter4.tex
Some misprints have been corrected

@@
 be considered.
-Let $\varphi(x)$ and $g_j(x)\le 0,1\le j\le m$ be the real functions defined in a closed interval $[a,b]$ of the real axis.
+Let $\varphi(x)$ and $g_j(x)\le 0,1\le j\le m$, be the real functions defined in a closed interval $[a,b]$ of the real axis.
 It is requires to find the values $x^*$ satisfying the condition
@@
   \label{eq4:problem}
-  \varphi(x^*)=\min\{\varphi(x):x\in [a,b], g_j(x)\le 0,\; 1\le j\le m\}
-\end{equation}
-
-As before, the objective function $\varphi$ (hereinafter denoted also as $g_{m+1}$) and the left parts of the constraints $g_j(x)\le 0,1\le j\le m$ re assumed to satisfy \emph{Lipschitz condition} with the corresponding constants $L_j,1\le j\le m+1$, namely:
+  \varphi(x^*)=\min\{\varphi(x):x\in [a,b], g_j(x)\le 0,\; 1\le j\le m\}.
+\end{equation}
+
+As before, the objective function $\varphi$ (hereinafter denoted also as $g_{m+1}$) and the left parts of the constraints $g_j(x)\le 0,1\le j\le m$, are assumed to satisfy \emph{Lipschitz condition} with the corresponding constants $L_j,1\le j\le m+1$, namely:
 \begin{equation}
   \label{eq4:lip_condition}
-  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2\in [a,b]
+  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2\in [a,b].
 \end{equation}
@@
 \label{ex4:problem}
-Let us consider a problem of type \eqref{eq4:problem} where $x\in [0.6;2.2],m=2$,
+Let us consider a problem of type \eqref{eq4:problem} where $x\in [0.6,2.2],\; m=2$,
 \begin{gather*}
@@
 \end{gather*}
-he exact solution of the problem is $x^*=2.0795,\varphi(x^*)=0.565$. The graphs of the functions $g_1(x),g_2(x)$,and $\varphi(x)$ are presented in Fig. \ref{fig:4_1}.
+The exact solution of the problem is $x^*=2.0795,\; \varphi(x^*)=0.565$. The graphs of the functions $g_1(x),g_2(x)$ and $\varphi(x)$ are presented in Fig. \ref{fig:4_1}.
 \begin{figure}[ht]
@@
 \section{Partial computability and index scheme of the account of constraints}
-Problem \eqref{eq4:problem} can be considered in the statement when each function $g_j,1\le j\le m+1$ is defined and computable only in the corresponding subdomains $Q_j\subset [a,b]$ where
+Problem \eqref{eq4:problem} can be considered in the statement when each function $g_j,1\le j\le m+1$, is defined and computable only in the corresponding subdomains $Q_j\subset [a,b]$, where
 \begin{equation}
@@
   \label{eq4:condition}
-  g_j(x)\le 0,1\le j \le \nu-1,g_\nu(x)>0,
+  g_j(x)\le 0,\; 1\le j \le \nu-1,\; g_\nu(x)>0,
 \end{equation}
@@
 \section{The index global search algorithm}
-The first trial is executed at an arbitrary internal point $x_1\in(a,b)$. The selection of any subsequent trial point $x^{k+1},k\ge 1$ is determined by the following rules.
+The first trial is executed at an arbitrary internal point $x_1\in(a,b)$. The selection of any subsequent trial point $x^{k+1},\; k\ge 1$ is determined by the following rules.
 
@@
   \[
-  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2 \in[a,b]
+  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,\; 1\le j\le m+1,\; x_1,x_2 \in[a,b]
   \]
@@
   \[
-  r_\nu\mu_\nu > 2L_\nu,1\le\nu\\le m+1.
+  r_\nu\mu_\nu > 2L_\nu,\; 1\le\nu \leq m+1.
   \]
@@
 
-n the cases, when it is necessary to denote the parameters $m$ and $\delta$ used by the generator during its functioning, let us use the notation $G ( m , \delta )$.
+In the cases, when it is necessary to denote the parameters $m$ and $\delta$ used by the generator during its functioning, let us use the notation $G ( m , \delta )$.
 In the experiments, the results of whichare presented below two generators were employed.
@@
 \begin{displaymath}
-  \varphi(x)=-\sum_{j=1}^{10}\frac{1}{(K_j(x-A_j)^2+C_j)}x\in[0,10].
+  \varphi(x)=-\sum_{j=1}^{10}\frac{1}{(K_j(x-A_j)^2+C_j)},\; x\in[0,10].
 \end{displaymath}

### 2017-10-05 2017/ParOptBook/Chapter4.tex
parOptBook: update refs in Chapter4

@@
 %\usepackage{cite}
+%\usepackage{amsmath}
+%\usepackage{amssymb}
+%
 %% see the list of further useful packages
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %\begin{document}
@@
 \begin{equation}
+  \label{eq4:13}
   I_\nu =\left\{i:1 \leq i \leq k, \; \nu=\nu(x_i) \right\}, \; 1 \leq \nu \leq m+1,
@@
 \begin{gather}
+  R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}), \nonumber \\
   \label{eq4:characteristic}
-  R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}), \nonumber \\
   R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),\\
@@
     \begin{equation}
+      \label{eq4:27}
       0< \epsilon_\nu <|g_\nu (x^* )|;
@@
     \item the convergence to the limit point $x$ is a bilateral one if $x\not=a$ and $x\not=b$.
-    If the conditions \eqref{eq4:reserves} and (4.27) are not satisfied for given vector of reserves $\epsilon_R$ but the
+    If the conditions \eqref{eq4:reserves} and \eqref{eq4:27} are not satisfied for given vector of reserves $\epsilon_R$ but the
     problem $\eqref{eq4:problem}$ has an $\epsilon$-reserved solution $x_\epsilon$ defined by \eqref{eq4:eps_res_problem}, then:
@@
 These difficulties are possible to avoid by
-substitution of the unknown coefficients $L_j,\: 1\le j\le m$, by their adaptive underestimates $\mu_\nu$ from (4.15) and by assumption that $\beta-\alpha>\delta=\epsilon q$ where $\epsilon >0$ (accuracy) is a parameter defined in the termination condition and the factor $q>1$ can be interpreted as an a priori estimate of how many times the length of the interval $[\alpha,\beta]$ is greater than given accuracy $\epsilon$.
+substitution of the unknown coefficients $L_j,\: 1\le j\le m$, by their adaptive underestimates $\mu_\nu$ from \eqref{eq4:lip_est} and by assumption that $\beta-\alpha>\delta=\epsilon q$ where $\epsilon >0$ (accuracy) is a parameter defined in the termination condition and the factor $q>1$ can be interpreted as an a priori estimate of how many times the length of the interval $[\alpha,\beta]$ is greater than given accuracy $\epsilon$.
 As a result, one can take
@@
 
-The problems were solved by the initial index algorithm as well as by the algorithm with $\epsilon$-reservation. The parameters $r_\nu ,\: 1 \le \nu \le 6$, from (4.17) were chosen as
+The problems were solved by the initial index algorithm as well as by the algorithm with $\epsilon$-reservation.
+The parameters $r_\nu ,\: 1 \le \nu \le 6$, from \eqref{eq4:characteristic} were chosen as
 \begin{displaymath}
@@
 Assume that every particular point $x$ from the search domain $[a,b]$ can have its own order of the problem constraints’ checking.
-The order of such checking corresponding to the point $x_i$ from (4.12) will be defines as some permutation $H(x_i )$ of the basic numbers from \eqref{eq4:basic_h}, i.e.,
+The order of such checking corresponding to the point $x_i$ from \eqref{eq4:characteristic} will be defines as some permutation $H(x_i )$ of the basic numbers from \eqref{eq4:basic_h}, i.e.,
 \begin{equation}
@@
 Since the constraints’ check order while the trial execution at the point $x$ i is determined by the corresponding tuple $H(x_i )$ from (4.33), the termination of the trial at this point after the computation of the values of l problem functions produces the value $z_i=g_\nu(x_i)$ where $\nu=j_{il}$.
-In this connection, let us use not the serial number $l$ of the last function calculated in this point, as it was accepted in (4.7) but the index of the last function in the basic numeration (4.32) as the
+In this connection, let us use not the serial number $l$ of the last function calculated in this point, as it was accepted in \eqref{eq4:trial} but the index of the last function in the basic numeration (4.32) as the
 value of $\nu(x_i )$ for the point $x_i$ , i.e.,
@@
 Check Order Method} (ACOM).
-Since the alterable constraint check order excludes a possibility to use expression (4.13) to determine the sets $I\nu$, to construct them, an information array connecting each $x_i$ from (4.12) with the indices of all functions, which have been computed in this node is necessary.
+Since the alterable constraint check order excludes a possibility to use expression \eqref{fig:4_13} to determine the sets $I\nu$, to construct them, an information array connecting each $x_i$ from \eqref{eq4:search_data} with the indices of all functions, which have been computed in this node is necessary.
 Hereinafter, the existence of such database is presumed.
@@
 \emph{Modification 2 (the determination of the trial points’ indices).} Each point $x_i$ from \eqref{eq4:search_data} is assigned the value $z_i=g_\nu(x_i)$ where according to (4.35), the index $\nu=\nu(x_i)$ is the basic index from (4.32) corresponding to the value of the last function estimated in this point.
-These indices are used in all computations based on expressions (4.13)– (4.17), (4.19).
-
-The modifications in the method of using the indices are related to rules (4.17) of calculations of the characteristics $R(i)$ of the intervals $(x_{i-1},x_i),\: 1\le i\le k$, with noncoincident
+These indices are used in all computations based on expressions \eqref{fig:4_13}-\eqref{eq4:characteristic}, \eqref{eq4:next_point}.
+
+The modifications in the method of using the indices are related to rules \eqref{eq4:characteristic} of calculations of the characteristics $R(i)$ of the intervals $(x_{i-1},x_i),\: 1\le i\le k$, with noncoincident
 indices of the boundary points only. Assume
@@
 \end{equation}
-are fulfilled, in order to determine the characteristic $R(i)$ one should use the second expression from formulae (4.17) with $\nu=\nu(x_i)$ (note that in this case the value $g_\nu(x_{i-1})$ is undefined).
+are fulfilled, in order to determine the characteristic $R(i)$ one should use the second expression from formulae \eqref{eq4:characteristic} with $\nu=\nu(x_i)$ (note that in this case the value $g_\nu(x_{i-1})$ is undefined).
 
@@
 \end{equation}
-the characteristic $R(i)$ is defined by the third expression from (4.17) with $\nu=\nu(x_{i-1})$ (in this
+the characteristic $R(i)$ is defined by the third expression from \eqref{eq4:characteristic} with $\nu=\nu(x_{i-1})$ (in this
 case, the value $g_\nu(x_i ) is undefined)$.
@@
   where $\beta-\alpha$ is the length of the interval $[\alpha,\beta]$ belonging to the feasible domain $Q$ from \eqref{eq4:Q_union} and including the point $x^*$, which is the solution of the problem;
-  \item since some sufficiently large step number $k$ the values $\mu_\nu$ from (4.15) corresponding to the nonempty sets $I_\nu$ satisfy the inequalities
+  \item since some sufficiently large step number $k$ the values $\mu_\nu$ from \eqref{eq4:lip_est} corresponding to the nonempty sets $I_\nu$ satisfy the inequalities
   \begin{displaymath}
@@
 problem solving practically coincide to the ones marked by three rows of the vertical strokes in Fig.
-4.6.
+\ref{fig:4_6}.
 However, the number of the computations of the constraint values appears to be different for FCOM and ACOM.
@@
 The experiments have been carried out for the algorithms described above with the fixed constraints’ check order and with the adaptive one (FCOM and ACOM, respectively).
-The parameters $r_\nu, 1\le\nu\le 5$, from (4.17) were defined by the expressions
+The parameters $r_\nu, 1\le\nu\le 5$, from \eqref{eq4:characteristic} were defined by the expressions
 \begin{displaymath}
@@
 Moreover, the trials executed according to the rules of ACOM appear to be much less computation costly (in average) than the ones executed by FCOM.
-This fact is illustrated in figures 4.16 and 4.17, where the average number of computations of the function values executed by FCOM and ACOM for given sets are presented.
+This fact is illustrated in figures \ref{fig:4_16} and \ref{fig:4_17}, where the average number of computations of the function values executed by FCOM and ACOM for given sets are presented.
 Fig.
-4.16 corresponds to the generator $G_{SH}$ while Fig.
-4.17 shows the results for the generator $G_{HL}$.
+\ref{fig:4_16} corresponds to the generator $G_{SH}$ while Fig.
+\ref{fig:4_17} shows the results for the generator $G_{HL}$.
 In both figures the function indices are plotted on the abscissa axes.
@@
   \end{equation}
+  \label{eq4:4_61}
   Then, the property (4.56) is a consequence of inequality \eqref{eq4:4_60}.
@@
   \end{displaymath}
-  where $r$ is from (4.61).
+  where $r$ is from \eqref{eq4:4_61}.
 Thus, at any search step $k\ge 1$ there exists an interval having a strictly positive characteristic.
@@
   \begin{gather}
+    \label{eq4:64}
     z_{t-1}\le L_\nu(\alpha - x_{t-1})\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t-1}) \\
+    \label{eq4:65}
     z_t\le L_\nu(x_{t}-\beta)\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t})
@@
   \begin{equation}
+    \label{eq4:66}
     z_t+z_{t-1}\le L_\nu \Delta_t - L_\nu(\beta-\alpha).
@@
 
-  Now, from the characteristic rules \eqref{eq4:characteristic_par}, from inequalities (4.64)–(4.66), and from the inequality
+  Now, from the characteristic rules \eqref{eq4:characteristic_par}, from inequalities \eqref{eq4:64}–\eqref{eq4:66}, and from the inequality
   \begin{displaymath}
@@
 
-  From (4.58), \eqref{eq4:67}, and from the rule (4.48) determining the choice of the interval for the next trial point, the inevitability of carrying out a trial in the interval $[x_{t-1},x_t]$ from (4.63) follows.
+  From (4.58), \eqref{eq4:67}, and from the rule (4.48) determining the choice of the interval for the next trial point, the inevitability of carrying out a trial in the interval $[x_{t-1},x_t]$ from \eqref{eq4:63} follows.
 Consequently, the interval $[x_{t-1},x_t]$ will be divided into two subintervals.
@@
 
-  Since the coincidence $x^*$ with one of the trial points means (4.62) to hold, then it remains to consider the case when $x^*\in(x_{t- 1} ,x_t ),\: t=t(k)$ for any $k\ge 1$.
+  Since the coincidence $x^*$ with one of the trial points means \eqref{eq4:62} to hold, then it remains to consider the case when $x^*\in(x_{t- 1} ,x_t ),\: t=t(k)$ for any $k\ge 1$.
 Here, because of the first condition of Theorem \ref{th4:convergence_par} and the definition (4.31) either $[x_{t-1},x^*]\in Q$ or $[x^* ,x_t]\in Q$ if the number $k$ is sufficiently large.
@@
   \begin{equation}
+    \label{eq4:70}
     [x_{t-1},x_t]\subset(u,w),\:t=t(q^s),\:s\ge S.
@@
   \begin{equation}
+    \label{eq4:71}
     g_j(x)>0,x\in D_j,\:1\le j\le m.
@@
   \begin{equation}
+    \label{eq4:72}
     (u,\overline x]\subset \cap_{j\in M}D_j,\:[\overline x,w)\subset \cap_{j\in N},
@@
 
-  According to the trial execution rules and to conditions (4.35) and (4.72), for any interval from (4.70) the relations
+  According to the trial execution rules and to conditions (4.35) and \eqref{eq4:72}, for any interval from \eqref{eq4:70} the relations
   \begin{displaymath}
@@
 
-  Now, from (4.46), (4.74), and (4.57) one can conclude that if the point $x$ is not a feasible one, it cannot be a limit point of the sequence $\{x^k\}$, i.e., for any limit point $x$ the condition \eqref{eq4:68} is true.
+  Assume $s>1$ to be so large that for $s>S$ the inequality $q^s\ge h$ where $h$ is from \eqref{eq4:65} holds. Additionally, according to (4.46),
+  \begin{equation}
+    \label{eq4:73}
+    z^*_\nu=-\epsilon_\nu,\:1\le\nu\le m.
+  \end{equation}
+  Then, from (4.47), (4.56), \eqref{eq4:71}, and \eqref{eq4:73} we get the estimate
+  \begin{equation}
+    \label{eq4:74}
+    R(t(q^p))<2\Delta_t-4\epsilon_\nu(r_\nu L_\nu)<0,\:p\ge s,
+  \end{equation}
+  where $\nu$ is an index from the set $\{\omega_1,\omega_2\}$.
+  Now, from (4.46), \eqref{eq4:74}, and (4.57) one can conclude that if the point $x$ is not a feasible one, it cannot be a limit point of the sequence $\{x^k\}$, i.e., for any limit point $x$ the condition \eqref{eq4:68} is true.
 \end{proof}

### 2017-10-05 2017/LION_Strongin/LION2017_BarkalovStrongin.tex
Unnecessary bracket has been deleted

@@
 The number of points in the additional grid $(\ref{Eq:6})$ for solving the least square problem (\ref{Eq:7}) was set at $20$.
-The problem of one-dimensional maximization and minimization from (\ref{Eq:9}) were solved by a scanning over a uniform grid of $100$ nodes within the interval $[a,b]$).
+The problem of one-dimensional maximization and minimization from (\ref{Eq:9}) were solved by a scanning over a uniform grid of $100$ nodes within the interval $[a,b]$.
 

### 2017-10-06 2017/LION_Barkalov_Lebedev/LION2017_BarkalovLebedev.tex
Some references have been corrected

@@
 The difficulty of a class was increased either by decreasing the radius of the attraction region of the global minimizer, or by decreasing the distance from the global minimizer $y^\ast$ to the domain boundaries.
-Application of this generator for studying some unconstrained optimization algorithms has been described in \cite{Kvasov2006}--\cite{SergeyevKvasov2015}.
+Application of this generator for studying some unconstrained optimization algorithms has been described in \cite{Grishagin2016}, \cite{Kvasov2006}--\cite{SergeyevKvasov2015}.
 
@@
 \bibitem{GergelLION} 
-Gergel, V.: An Approach for Generating Test Problems of Constrained Global Optimization. Proceedings of the LION11 conference (2017) [submitted]
+Gergel, V.: An approach for generating test problems of constrained global optimization.
+In: Battiti, R., Kvasov, D., Sergeyev, Y.
+(eds.) LION 2017.
+LNCS, vol.
+10556.
+Springer (2017) DOI:10.1007/978-3-319-69404-7\_24
 

### 2017-10-06 2017/ParOptBook/Chapter1.tex
Chapter 6 is ready
Some misprints have been corrected
Chapters 3 and 5 have been included. There is a question about cross-references between chapters.

@@
 Convergence of optimization methods and possibility of obtaining estimations of the problem solution after a finite number of objective function evaluations are discussed.
-Significant complexity of multidimensional multie
-xtremal optimization problems is considered and approaches of overcoming this complexity via reduction schemes are briefly described. }
+Significant complexity of multidimensional multiextremal optimization problems is considered and approaches of overcoming this complexity via reduction schemes are briefly described.
+}
 
 \section{The global optimization problem}
-\label{sec:1}
+\label{sec:1_1}
 The aim of this section is to introduce the necessary terms and to formulate the general problem of the global optimization, the investigation of which the present monograph is dedicated to. 
@@
 \begin{displaymath}
-\label{eq:12}
+%\label{eq:12}
 \varphi(y_1,y_2)=-1.5y_1^2\exp(1-y_1^2-20.25(y_1-y_2)^2)-
@@
 \begin{displaymath}
-\label{eq:12}
+%\label{eq:12}
 -(0.5(y_1-1)(y_2-1))^4\exp(2-(0.5(y_1-1))^4-(y_2-1)^4)
@@
 Thus, the feasible domain is disconnect and consists of three non-convex subdomains (the feasible domain is dark in Fig.~\ref{1_fig_1}).
-The global minimum $\varphi(y_1^*,y_2^*)=-1.489$  is achieved at the point $(y_1^*,y_2^*)$  = (0,942, 0,944).
+The global minimum $\varphi(y_1^*,y_2^*)=-1.489$  is achieved at the point $(y_1^*,y_2^*)$  = (0.942, 0.944).
 
 \section{Numerical optimization methods}
-\label{sec:1.2}
+\label{sec:1_2}
 
@@
 \begin{displaymath}
-\label{eq:19}
+%\label{eq:19}
 y_k^*=\arg \min_{1\leq i\leq k} \varphi(y^i).
@@
 \section{Multiextremal problems and reduction of complexity}
-\label{sec:3}
+\label{sec:1_3}
 In multiextremal problems the dimension affects the difficulty of solving such problems significantly.
@@
 \label{eq:31}
-\lim_{\tau \to \infty}\min_{1\leq k \leq \tau}=\varphi^*.
+\lim_{\tau \to \infty}\min_{1\leq k \leq \tau}{\varphi(y^k)}=\varphi^*.
 \end{equation}
-In other words, the trial sequences of these methods are everywhere dense in the search domain, i.e., converge in the sense of Definition 1.4 to all points within this domain and, consequently, to the global minimum point as well.
+In other words, the trial sequences of these methods are everywhere dense in the search domain, i.e., converge in the sense of Definition \ref{def:1_4} to all points within this domain and, consequently, to the global minimum point as well.
 
@@
 \begin{thebibliography}{99.}
-\bibitem{BetroSchoen} Betro, B., Schoen, F.: Sequential stopping rules for the multistart algorithm in global optimisation. Math. Program. \textbf{38}(3), 271-286 (1987)
-\bibitem{BoendeRinnooy}	Boender, C.G.E, Rinnooy Kan, A.H.G.: Bayesian stopping rules for multistart global optimization methods. Math. Program. \textbf{37}(1), 59-80 (1987)
-\bibitem{1_Butz}	Butz, A.R. Space-Filling Curves and Mathematical Programming.: Inform. Control \textbf{12}, 314-330 (1968)
-\bibitem{1_CarrHowe}	Carr, C.R., Howe, C.W. Quantitative Decision Procedures in Management and Economic: Deterministic Theory and Applications. McGraw-Hill, New York (1964)
+\bibitem{BetroSchoen} Betro, B., Schoen, F.: Sequential stopping rules for the multistart algorithm in global optimisation. Math. Program. \textbf{38}(3), 271--286 (1987)
+\bibitem{BoendeRinnooy}	Boender, C.G.E, Rinnooy Kan, A.H.G.: Bayesian stopping rules for multistart global optimization methods. Math. Program. \textbf{37}(1), 59--80 (1987)
+\bibitem{1_Butz}	Butz, A.R. Space-Filling Curves and Mathematical Programming.: Inform. Control \textbf{12}, 314--330 (1968)
+\bibitem{1_CarrHowe}	Carr, C.R., Howe, C.W.: Quantitative Decision Procedures in Management and Economic: Deterministic Theory and Applications. McGraw-Hill, New York (1964)
 \bibitem{1_Evtushenko}	Evtushenko, Yu.G.: Numerical Optimization Techniques. Translation Series in Mathematics and Engineering. Optimization Software  Inc., Publication Division, New York (1985)
 \bibitem{1_FloudasPardalos} Floudas, C.A., Pardalos, P.M.: State of the Art in Global Optimization. Kluwer Academic Publishers, Dordrecht (1996)
-\bibitem{1_GerGriIsr} Gergel, V.,  Grishagin, V., Israfilov, R.: Local tuning in nested scheme of global optimization, Procedia Computer Science \textbf{51}, 865-874 (2015) 
+\bibitem{1_GerGriIsr} Gergel, V.,  Grishagin, V., Israfilov, R.: Local tuning in nested scheme of global optimization, Procedia Computer Science \textbf{51}, 865--874 (2015) 
 \bibitem{1_GerGriGer}	Gergel, V., Grishagin, V., Gergel, A.: Adaptive nested optimization scheme for multidimensional global search, J. Glob. Opt. \textbf{66}, 35–51 (2016).
-\bibitem{1_Goertzel}	Goertzel, B.: Global Optimization with Space-Filling Curves. Appl. Math. Lett. 12, 133-135 (1999)
+\bibitem{1_Goertzel}	Goertzel, B.: Global Optimization with Space-Filling Curves. Appl. Math. Lett. 12, 133--135 (1999)
 \bibitem{1_GriIsrAIP}	Grishagin, V.A., Israfilov, R.A.: Global search acceleration in the nested optimization scheme. AIP Conf. Proc. \textbf{1738}, 400010 (2016)
@@
 (2017) doi:10.1016/j.amc.2017.06.036
-\bibitem{1_GriIsrCEUR}Grishagin, V.A., Israfilov, R.A. Multidimensional Constrained Global Optimization in Domains with Computable Boundaries. CEUR Workshop Proceedings 1513, 75-84 (2015)
+\bibitem{1_GriIsrCEUR}Grishagin, V.A., Israfilov, R.A.: Multidimensional Constrained Global Optimization in Domains with Computable Boundaries. CEUR Workshop Proceedings 1513, 75--84 (2015)
 \bibitem{1_HimOliPet}	Hime, A.E., Oliveira Jr., H.A., Petraglia, A.: Global Optimization Using Space-Filling Curves and Measure-Preserving Transformations.
-Soft Computing in Industrial Applications \textbf{96}, 121-130 (2011)
+Soft Computing in Industrial Applications \textbf{96}, 121--130 (2011)
 \bibitem{1_Himmelblau}	Himmelblau, D.M.: Applied Nonlinear Programming. McGraw-Hill, New York (1972)
@@
 D.
-Journal of Basic Engineering \textbf{86}, 97-106 (1964)
+Journal of Basic Engineering \textbf{86}, 97--106 (1964)
 \bibitem{1_LeraSergCNSNS}	Lera, D., Sergeyev, Y.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and H{\"o}lder constants.
-Communications in Nonlinear Science and Numerical Simulation \textbf{23}, 328-342 (2015)
-\bibitem{1_LeraSergANM}	Lera, D., Sergeyev, Y.D.: Lipschitz and H{\"o}lder global optimization using space-filling curves. Applied Numerical Mathematics \textbf{60}, 115-129 (2010) 
+Communications in Nonlinear Science and Numerical Simulation \textbf{23}, 328--342 (2015)
+\bibitem{1_LeraSergANM}	Lera, D., Sergeyev, Y.D.: Lipschitz and H{\"o}lder global optimization using space-filling curves. Applied Numerical Mathematics \textbf{60}, 115--129 (2010) 
 \bibitem{1_McCormick}	McCormick, G.P.: Nonlinear Programming: Theory, Algorithms and Applications. John Wiley and Sons, New York (1988)
@@
 \bibitem{1_Pinter}	Pint{\'e}r, J.D.: Global Optimization in Action.: Kluwer Academic Publishers, Dordrecht (1996) 
-\bibitem{1_Piyavskij}	Piyavskij, S.A.: An Algorithm for Finding the Absolute Extremum of a Function. USSR Comput. Math. Math. Phys. \textbf{12}(4), 57-67 (1972)
-\bibitem{1_SergGriJCAA}	Sergeyev, Y.D., Grishagin, V.A.: Parallel Asynchronous Global Search and the Nested Optimization Scheme. J. Comp. Analysis Appl. \textbf{3}, 123-145 (2001)
+\bibitem{1_Piyavskij}	Piyavskij, S.A.: An Algorithm for Finding the Absolute Extremum of a Function. USSR Comput. Math. Math. Phys. \textbf{12}(4), 57--67 (1972)
+\bibitem{1_SergGriJCAA}	Sergeyev, Y.D., Grishagin, V.A.: Parallel Asynchronous Global Search and the Nested Optimization Scheme. J. Comp. Analysis Appl. \textbf{3}, 123--145 (2001)
 \bibitem{1_SergKvasMonogr}	Sergeyev, Y.D., Kvasov, D.E.: Deterministic Global Optimization: Diagonal Approach Briefly. Springer-Verlag, New York (2017)
 \bibitem{1_SergStrLeraMonogr}	Sergeyev, Y.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting Space-Filling Curves. Springer (2013)
-\bibitem{1_ShiOlaf}   Shi, L., {\'O}lafsson, S.: Nested Partitions Method for Global Optimization. Operations Research \textbf{48}, 390-407 (2000)
+\bibitem{1_ShiOlaf}   Shi, L., {\'O}lafsson, S.: Nested Partitions Method for Global Optimization. Operations Research \textbf{48}, 390--407 (2000)
 \bibitem{1_StrMonRus}	Strongin, R.G.: Numerical Methods in Multiextremal Problems (Information-Statistical Algorithms). Nauka, Moscow (1978). In Russian
@@
 \bibitem{1_Sukharev1989}	Sukharev, A.G.: Minimax Algorithms in Problems of Numerical Analysis. Nauka, Moscow (1989). In Russian
-\bibitem{1_vanDam}	van Dam, E.R., Husslage, B., Hertog, D. One-dimensional Nested Maximin Designs. J. Glob. Opt. \textbf{46}, 287-306 (2010)
+\bibitem{1_vanDam}	van Dam, E.R., Husslage, B., Hertog, D. One-dimensional Nested Maximin Designs. J. Glob. Opt. \textbf{46}, 287--306 (2010)
 \bibitem{1_Vasiliev}	Vasiliev, F.P. Numerical Methods for Solving Extremal Problems. Nauka, Moscow (1988). In Russian
-\bibitem{1_ZhigZhil} Zhigljavsky, A.A., Zilinskas, A.: Stochastic Global Optimization. Springer, New York. (2008)
-\bibitem{1_Zielinski}	Zieli{\'n}ski, R.: A statistical estimate of the structure of multi-extremal problems. Math. Program. \textbf{21}(1), 348-356 (1981)
+\bibitem{1_ZhigZhil} Zhigljavsky, A.A., $\check{Z}$ilinskas, A.: Stochastic Global Optimization. Springer, New York. (2008)
+\bibitem{1_Zielinski}	Zieli{\'n}ski, R.: A statistical estimate of the structure of multi-extremal problems. Math. Program. \textbf{21}(1), 348--356 (1981)
 

### 2017-10-06 2017/ParOptBook/Chapter3.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter4.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter5.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter6.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter7.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter8.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\dedic.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\editor.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\foreword.tex
D:\Common\Barkalov\Публикации\20

--- 2017-09-01 2017/ParOptBook/Chapter7.tex
@@
-%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% sample root file for your "contribution" to a contributed volume
-%
-% Use this file as a template for your own input.
-%
-%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %\documentclass[graybox]{svmult}
-%
-%% choose options for [] as required from the list
-%% in the Reference Guide
 %
@@
 %\usepackage{type1cm}        % activate if the above 3 fonts are
-                             % not available on your system
-%
+                            %% not available on your system
+%%
 %\usepackage{makeidx}         % allows index generation
 %\usepackage{graphicx}        % standard LaTeX graphics tool
-%                             % when including figure files
+                             %% when including figure files
 %\usepackage{multicol}        % used for the two-column index
-%\usepackage[bottom]{footmisc}% places footnotes at page bottom
-%
-%% see the list of further useful packages
-%% in the Reference Guide
+%\usepackage[bottom]{footmisc}
+%\usepackage{amsmath}
+%\usepackage{multirow}
+%\usepackage{array}
+%\usepackage{caption}
 %
 %\makeindex             % used for the subject index
-%                       % please use the style svind.ist with
-%                       % your makeindex program
+                       %% please use the style svind.ist with
+                       %% your makeindex program
 %
@@
 
-\title{Software Systems for Global Optimization }
+\title{Parallel algorithms for one-dimensional multiextremal optimization}
+%\titlerunning{Global optimization via dimensionality reduction}
 % Use \titlerunning{Short Title} for an abbreviated version of
 % your contribution title if the original one is too long
-\author{Name of First Author and Name of Second Author}
+\author{Vladimir A.Grishagin and Yaroslav D.Sergeyev}
 % Use \authorrunning{Short Title} for an abbreviated version of
 % your contribution title if the original one is too long
-\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
-\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
-%
-% Use the package "url.sty" to avoid
-% problems with special characters
-% used in your e-mail or web address
+\institute{Vladimir A.Grishagin \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia   \email{vagris@unn.ru}
+\and Yaroslav D.Sergeyev \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia 
+\at DIMES, University of Calabria, Via P. Bucci, Cubo 42-C, 87036 Rende (CS), Italy \email{yaro@si.deis.unical.it}}
 %
@@
 
-\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\abstract{This chapter is devoted to the development and study of parallel global search algorithms on the base of a generalization of the sequential characteristical scheme to the case of parallel computations.
+Classes of synchronous and asynchronous optimization algorithms are introduced following the general schemes of parallel computations for solving optimization problems described in Chapter 1.
+The problems of convergence and parallelization efficiency for these characteristical algorithms are theoretically investigated including a study of various types of convergence, sufficient conditions of convergence to global minimizers, and conditions of non-redundant parallelization.
+The general theoretical results are applied to a number of known global optimization methods and are illustrated in computational experiments.}
 
-\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\section{Parallel synchronous characteristical algorithms of global search}
+\label{sec:3_1}
+Let us consider the one-dimensional problem 
+\begin{equation}
+\label{eq:3_1}
+\varphi(x)\rightarrow \min,\ x\in Q=[a,b]
+\end{equation}
 
-\section{Section Heading}
-Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.
+for the search domain  $Q=[a,b]$ and the general schemes of the parallel synchronous optimization methods (\ref{eq:1_20})-(\ref{eq:1_25}) and the parallel asynchronous methods (\ref{eq:1_26})-(\ref{eq:1_30}).
+Let us assume that the termination condition is absent in these schemes, i.e., in (\ref{eq:1_25}) $H_{k(n)}(\bullet) = 1$ and in (\ref{eq:1_30}) $H_{k}(\bullet)=1$ for any $k=k(n)$ will be the total number of trials executed during all $n$ iterations.
+
+\begin{definition}
+	A parallel method for solving the problem (\ref{eq:3_1}) is called \textit{parallel synchronous characteristical algorithm}, if, starting from some iteration number $n_0\ge1$, the selection of the trial coordinates at the $(n+1)$-th iteration $(n\ge n_0)$ is performed according to the following procedure:
+
+	\begin{description} [Step 1.]
+	\item [\textbf{Step 1.}] {Define a set 
+	\begin{equation}
+	\label{eq:3_2}
+	\Lambda_n = \{x_0,x_1,...,x_\tau \}
+	\end{equation}
+	of a finite number $\tau + 1 = \tau(n)+1$ of the points in the domain $Q=[a,b]$ assuming that $a \in \Lambda_n, b \in \Lambda_n$, all coordinates of the preceding trials $x^i \in \Lambda_n, 1 \le i \le k$, and the set $\Lambda_n$ is ordered (by the subscript) in the increasing order of the coordinates, i.e.,
+	\begin{equation}
+	\label{eq:3_3}
+	a = x_0 < x_1 < ... <x_{\tau - 1} < x_\tau = b.
+	\end{equation}
+	}
+	\item [\textbf{Step 2.}] {Juxtapose the value $R(i)$ called the characteristic of the interval to each interval $(x_{i-1}, x_i), 1 \le i \le \tau$.}
+
+	\item [\textbf{Step 3.}] {Sort the characteristics $R(i), 1 \le i \le \tau$, in the decreasing order:
+	\begin{equation}
+	\label{eq:3_4}
+	R(t_1) \ge R(t_2) \ge R(t_{\tau - 1}) \ge R(t_\tau).
+	\end{equation}
+	}
+	\item [\textbf{Step 4.}] {Select $p = p(n+1) \le \tau$ first characteristics from the series (3.4) with the numbers $t_j, 1 \le j \le p$, and to choose the new trials of the next $(n+1)_{th}$ iteration at the points of the set
+	\begin{equation}
+	\label{eq:3_5}
+	T(n+1) = \{x^{k+1}, ...,x^{k+p}\},
+	\end{equation}
+	according to the rule
+	\begin{equation}
+	\label{eq:3_6}
+	x^{k+j} = d(t_j) \in (x_{t_j - 1}, x_{t_j})
+	\end{equation}
+	for $1 \le j \le p$, in the intervals with maximal characteristics.}
+	\end{description}
+\end{definition}
+
+As the examples of the parallel characteristical algorithms one can consider the parallel versions of the scanning method, of the  method of broken lines, of Kushner’s method, and of the information algorithms, the sequential variants of which have been described in the previous Chapter.
+
+The class of the parallel characteristical methods has been introduced and investigated in (\cite{3_GrishaginSergeyevStrongin}).
+In this work, the succession of the convergence properties of the parallel characteristical methods with respect to the sequential prototypes has been established.
+Also, the estimates of the speed up of the parallel characteristical algorithms have been obtained.
+A theoretical substantiation of the properties of the synchronous parallel characteristic methods with respect to the character of their convergence and the parallelization efficiency will be provided below.
+
+\section{Convergence of characteristical algorithms}
+\label{sec:3_2}
+\subsection{Bilateral convergence}
+Let us introduce the notations $z_j = \varphi(x_j), 1 \le j \le \tau$, for the values of the objective function at the points $x_j \in \Lambda_k$ and introduce the values
+\begin{equation*}
+\xi = min \{ z_{i-1}, z_i \},
+\end{equation*}
+\begin{equation*}
+\tilde{\xi_i} = 
+\begin{cases}
+\xi_i, & if\  x_{i-1} \in \{x^S\}\  and\  x_i \in \{x^S\}; \\
+z_{i-1},& if\  x_{i-1} \in \{x^S\}\  and \ x_i \notin \{x^S\}; \\
+z_i, & if\  x_{i-1} \notin \{x^S\}\  and\  x_i \in \{x^S\}.
+\end{cases}
+\end{equation*}
+
+\begin{theorem}
+\label{theor:3_1}
+	Let a point $x_\ast$ be a limit point (or an accumulation point) of the search trials sequence $\{x_k\}$  generated by a synchronous characteristical algorithm when solving the problem (3.1) over the interval $[a, b]$, and $x_\ast \neq a$  and $x_\ast \neq b $.
+Assume that the characteristics $R(i)$ and the rules for the selection of the next trial points $d(t)$ satisfy the following requirements:
+
+	\textbf{i)} if a point $\bar{x} \in [x_{i(n)-1}, x_{i(n)}]$ and $x_{i(n)-1} \to \bar{x}, x_{i(n)} \to \bar{x}$ when $n \to \infty $  then
+	\begin{equation}
+	\label{eq:3_7}
+	R(i(n)) \to - \mu \varphi(\bar{x}) + c;
+	\end{equation}
+
+	\textbf{ii)} in the case, when starting from some search iteration, the interval $(x_{i-1}, x_i), i=i(n)$, does not include trial points, i.e., a number $\tilde{n} \ge 1$ exists such that for any $n \ge \tilde{n}$
+	\begin{equation}
+	\label{eq:3_8}
+	(x_{i-1} ,x_{i} )\bigcap \{ x^{k} \} =\emptyset
+	\end{equation}
+	for the characteristic of the interval holds the strict inequality
+	\begin{equation}
+	\label{eq:3_9}
+	\mathop{\lim }\limits_{n\to \infty } R(i)>-\mu \tilde{\xi }_{i} +c
+	\end{equation}
+
+	\textbf{iii)} when selecting a point for the next trial from set (2.59) the inequality 
+	\begin{equation}
+	\label{eq:3_10}
+	{\rm \; \; }\max \{ x^{k+j} -x_{t_{j} -1} ,x_{t_{j} } -x^{k+j} \} \le \nu (x_{t_{j} } -x_{t_{j} -1} )
+	\end{equation}
+	takes place where $\mu$, $c$, and $\nu$ are some constants, and $\mu \ge 0,{\rm \; \; }0<\nu <1$.
+
+	Then the sequence $\{ x^{k} \} $ includes two subsequences; one of those converges to $x^{*} $ from the left, and another one from the right.
+	\end{theorem}
+	\begin{proof} Let us first consider the case when $x^{*} \notin \{ x^{k} \} $.
+Let us denote as $s=s(n),n\ge 1$, the number of the interval $(x_{s-1} ,x_{s} )$ including the limit point $x^{*} $ after the execution of the $n$-th iteration.
+Obviously, after the first iteration $[x_{s-1} ,x_{s} ]\subseteq [a,b]$.
+After the next trial point $x^{k+l} \in T(n+1)$ has fallen into the interval $(x_{s-1} ,x_{s} )$, the following estimate holds according to (3.10) for a new interval $(x_{s(n+1)-1} ,x_{s(n+1)} )$ containing $x^{*} $:
+
+	\begin{equation*}
+		x_{s(n+1)} -x_{s(n+1)-1} \le \nu (x_{s(n)} -x_{s(n)-1} )
+	\end{equation*}
+
+	In this case, if $\lambda$ trials (starting from the beginning of the search process) fell into the interval including the point $x^{*} $, the length of this interval satisfies the inequality 
+	\begin{equation}
+	\label{eq:3_11}
+	x_{s} -x_{s-1} \le \nu ^{\lambda } (b-a)
+	\end{equation}
+
+	Since the point $x^{*}$ is a limit one, after the formation of the interval $(x_{s-1} ,x_{s} )$ at some iteration, an infinite number of trials falls into this interval.
+Therefore, it follows from (\ref{eq:3_11}) that
+	\begin{equation}
+	\label{eq:3_12}
+	\mathop{\lim }\limits_{n\to \infty } (x_{s(n)} -x_{s(n)-1} )=0 
+	\end{equation}
+
+	According to  (\ref{eq:3_12}), we can take the sequence of the left ends $\{ x_{s(n)-1} \} $ and the one of the right ends $\{ x_{s(n)} \} $ of intervals $(x_{s(n)} ,x_{s(n)-1} )$ as the sought subsequences.
+
+	If $x^{*} \in \{ x^{k} \} $, the numbers $w\ge 1$ and $q>k(w-1)$ such that $x^{q} =x^{*} \in T(w)$ can be always found.
+Then for any $n>w$ a number $j=j(n)$, $0\le j\le \tau (n)$, exists, for which $x_{j} =x^{*} $.
+Assume that a unilateral convergence to $x^{*} $ takes place, for example, from the left (the case of the unilateral convergence from the right can be considered in the same way).
+Then an iteration number $\tilde{w}\ge w$ and the trial number $v>k(\tilde{w})$ can be always found so that $x^{v} \in T(\tilde{w}),$ and for all $n\ge \tilde{w}$ the trials will not fall into the interval $(x_{j} ,x_{j+1} )=(x^{q} ,x^{v} )$.
+
+	For the interval $(x_{j} ,x_{j+1} )$ according to  (\ref{eq:3_8}),  (\ref{eq:3_9})
+
+	\begin{equation*}
+	\mathop{\lim }\limits_{n\to \infty } R(j+1)>-\mu \, \tilde{\xi }_{j+1} +c\ge -\mu \, \varphi (x^{*} )+c.
+	\end{equation*}
+
+	Correspondingly, for the interval $(x_{j-1} ,x_{j} )$ because of  (\ref{eq:3_7})
+
+	\begin{equation}
+	\label{eq:3_13}
+	\mathop{\lim }\limits_{n\to \infty } R(j)=-\mu \, \varphi (x^{*} )+c.
+	\end{equation}
+
+	Therefore, starting from some iteration, the inequality $R(j+1)>R(j)$ will hold.
+
+	However, because of the decision rules  (\ref{eq:3_4}) - (\ref{eq:3_6}) this relation contradicts to the impossibility to execute the trials in the interval $(x_{j} ,x_{j+1} )$. 
+	\end{proof}
+	\begin{corollary} 
+	A termination condition of the kind 
+		\begin{equation}
+	\label{eq:3_14}
+	\mathop{\min }\limits_{1\le j\le p} (x_{t_{j} } -x_{t_{j} -1} )\le \varepsilon ,
+	\end{equation}
+	where the numbers $t_{j} ,\; 1\le j\le p$, are from  (\ref{eq:3_4}) and $\varepsilon >0$ is the predefined search precision can be introduced into the computational scheme of the characteristical algorithm satisfying the conditions of Theorem \ref{theor:3_1}.
+In other words, the computations are terminated as soon as the length of the smallest interval among the intervals of the iteration with the maximal characteristics becomes less than the predefined precision $\epsilon$.
+In this case, the search process is terminated after a finite number of steps.
+	\end{corollary}
+	\begin{proof} First, note that in a finite interval $[a,b]$ the trial sequence will always have at least one limiting point $x^{*} $.
+Let us denote as $s=s(k),k\ge 1$, the number of the interval including the point $x^{*} $ at the $n$-th iteration of the search.
+Since this point is a limit one, an infinite number of trials falls into the interval $(x_{s-1} ,x_{s} )$.
+In this case, relation  (\ref{eq:3_12}) takes place for this interval, from  where it follows that condition  (\ref{eq:3_14}) is fulfilled inevitably at some search iteration.
+Note that if the point $x^{*} $ is not an internal one, a unilateral convergence is sufficient for feasibility of  (\ref{eq:3_14}).
+	\end{proof}
+	\textbf{Note.} Condition  (\ref{eq:3_10}) is equivalent to the inequality
+	\begin{equation*}
+	{\rm \; \; }\min \{ x^{k+j} -x_{t_{j} -1} ,x_{t_{j} } -x^{k+j} \} \ge (1-\nu )(x_{t_{j} } -x_{t_{j} -1} ).
+	\end{equation*}
+
+	Indeed, one can see easily that, for example,
+	\begin{displaymath}
+	x^{k+j} -x_{t_{j} -1} =(x_{t_{j} } -x_{t_{j} -1} )-(x_{t_{j} } -x^{k+j} )\ge (x_{t_{j} } -x_{t_{j} -1} )-\nu (x_{t_{j} } -x_{t_{j} -1} )=  
+	\end{displaymath}
+	\begin{displaymath}
+	=(1-\nu )(x_{t_{j} } -x_{t_{j} -1} ).  
+	\end{displaymath}
+
+	For the interval $x_{t_{j} } -x^{k+j} $, an analogous estimate takes place, from were it follows that the considered inequality is true.
+
+	Let us verify the feasibility of the conditions of Theorem  \ref{theor:3_1} for several examples of the algorithms belonging to the caharacteristical class.
+These algorithms are generalizations of known sequential methods of global optimization (see Chapter 2) to the parallel case.
+\\
+
+	\textbf{Method of sequential scanning } \\
+
+	For this method the characteristics 
+	\[R(i)=x_{i} -x_{i-1} ,\, 1\le i\le \tau ,\] 
+	and the new point  $x^{k+j} ,\; 1\le j\le p,$ from  (\ref{eq:3_6}) in the interval $(x_{t_{j} -1} ,x_{t_{j} } )$is calculated as
+	\[x^{k+j} =0.5(x_{t_{j} -1} +x_{t_{j} } ).\]
+
+	If the interval $(x_{i-1} ,x_{i} )$ shrinks into a point, the characteristic of this method tends to zero.
+Therefore, one can take the constants $\mu$ and $c$ in  (\ref{eq:3_7}) as $\mu = c = 0 $.
+If the trials do not fall into the interval $(x_{i-1} ,x_{i} )$ starting from some search iteration, the length of this interval (coinsiding with its characteristic) remains positive that provides the fulfillment of the condition  (\ref{eq:3_9}) for selected $\mu$ and $c$.
+As for inequality  (\ref{eq:3_10}), it obviously is true for  $\nu = 0.5$.
+	\\
+
+	\textbf{Kushner's method }\cite{3_Kushner} \\
+
+	In this algorithm the characteristics 
+	\[R(i)=-\frac{4(\varphi _{k}^{*} -\delta _{k} -z_{i} )(\varphi _{k}^{*} -\delta _{k} -z_{i-1} )}{x_{i} -x_{i-1} } ,\; 1\le i\le \tau ,\] 
+	are used, and the new trial  (\ref{eq:3_6}) in the interval  $(x_{t-1} ,x_{t} ),\, t=t_{j} ,$  are executed at the point
+	\begin{equation} 
+	\label{eq:3_15} 
+	x^{k+j} =x_{t-1} +\frac{(x_{t} -x_{t-1} )(\varphi _{k}^{*} -\delta _{k} -z_{t} )}{2(\varphi _{k}^{*} -\delta _{k} )-z_{t} -z_{t-1} }  
+	\end{equation} 
+	where $\delta _{k} >0$ is a parameter of the method and 
+	\[\varphi _{k}^{*} =\mathop{\min }\limits_{0\le i\le \tau } z_{i} .\]
+
+	Let us assume that the function \textit{$\varphi$}(\textit{x}) is limited within the finite values $\varphi _{\min } $ and $\varphi _{\max } $ over the interval $[a,b]$, i.e., $\varphi _{\min } \le \varphi (x)\le \varphi _{\max } ,\; x\in [a,\, b]$.
+
+	If for any step $k$ the method parameter $\delta _{k} >\delta >0$ then $\varphi _{k}^{*} -z_{i} -\delta _{k} <-\delta <0$, $\varphi _{k}^{*} -z_{i-1} -\delta _{k} <-\delta <0$ and the characteristic 
+	\[R(i(k))=-\frac{4(\varphi _{k}^{*} -\delta _{k} -z_{i} )(\varphi _{k}^{*} -\delta _{k} -z_{i-1} )}{x_{i} -x_{i-1} } \to -\infty \] 
+	when the length of the interval $(x_{i-1} ,x_{i} )$ tends to zero.
+Therefore,  (\ref{eq:3_7}) holds when $\mu = 0$ and the nominal value of $c=-\infty $ (the proof of Theorem \ref{theor:3_1} does not change if the limit value in the right part of  (\ref{eq:3_7}) is equal to $-\infty$).
+
+	Since the characteristic of any interval with a nonzero length possesses a finite value,  (\ref{eq:3_9}) holds as well.
+
+	Let us remind a notation $t=t_{j} $ for the trial $x^{k+j} $of the current iteration  (\ref{eq:3_15}) and consider the differences
+
+	\begin{equation*}
+	x^{k+j} -x_{t-1} =\beta _{1} (x_{t} -x_{t-1} ),\; \; x_{t} -x^{k+j} =\beta _{2} (x_{t} -x_{t-1} ), 
+	\end{equation*}
+
+	where
+
+	\begin{equation*}
+	\beta _{1} =\frac{z_{t-1} -\varphi _{k}^{*} +\delta _{k} }{z_{t-1} +z_{t} -2(\varphi _{k}^{*} -\delta _{k} )} ,\; \; \beta _{2} =\frac{z_{t} -\varphi _{k}^{*} +\delta _{k} }{z_{t-1} +z_{t} -2(\varphi _{k}^{*} -\delta _{k} )}.
+	\end{equation*}
+
+	Let us introduce an auxiliary function $\sigma (w,\alpha )=\frac{w+\alpha }{w+2\alpha } $.
+Since $\min \{ z_{t-1} ,\; z_{t} \} \ge \varphi _{k}^{*} $, $\beta _{1} \le \sigma (z_{t-1} -\varphi _{k}^{*} ,\delta _{k} ),\; \beta _{2} \le \sigma (z_{t} -\varphi _{k}^{*} ,\delta _{k} )$.
+
+	For any positive value $w$, the function $\sigma (w,\alpha )$ decreases with increasing $\alpha$ because $\sigma '_{\alpha } (w,\alpha )=\frac{-w}{(w+2\alpha )^{2} } <0,$ therefore, for $\delta _{k} >\delta >0$ the following inequalities hold:
+
+	\begin{equation*}
+	\beta _{1} \le \sigma (z_{t-1} -\varphi _{k}^{*} ,\delta ),\; \beta _{2} \le \sigma (z_{t} -\varphi _{k}^{*} ,\delta ).
+	\end{equation*}
+
+	In turn, the function $\sigma (w,\alpha )$ increases with increasing $w$ for any positive $\alpha$ since $\sigma '_{w} (w,\alpha )=\frac{\alpha }{(w+2\alpha )^{2} } >0$, therefore
+
+	\begin{equation*}
+	\max \{ \sigma (z_{t-1} -\varphi _{k}^{*} ,\delta ),\; \sigma (z_{t} -\varphi _{k}^{*} ,\delta )\} \le \sigma (\varphi _{\max } -\varphi _{\min } ,\; \alpha ). 
+	\end{equation*}
+
+	The last inequality means that
+
+	\begin{equation*}
+	\max \{ \beta _{1} ,\; \beta _{2} \} \le \sigma (\varphi _{\max } -\varphi _{\min } ,\; \delta )=\frac{\varphi _{\max } -\varphi _{\min } +\delta }{\varphi _{\max } -\varphi _{\min } +2\delta }. 
+	\end{equation*}
+
+	The last value can be selected as the constant $\nu$  for the relation  (\ref{eq:3_10}) since, obviously, $0<\nu <1$. \\
+
+	\textbf{Method of broken lines \cite{3_Piyavskij, 3_Shubert}} \\
+
+	The characteristics of this algorithm
+
+	\begin{equation}
+	\label{eq:3_16}
+	R(i)=\frac{m(x_{i} -x_{i-1} )}{2} -\frac{z_{i} +z_{i-1} }{2} ,\; 1\le i\le \tau,
+	\end{equation}
+	and for the interval $(x_{t_{j} -1} ,x_{t_{j} } )$ from  (\ref{eq:3_6}) the point  of the new trial falling into this interval is taken in the form
+
+	\begin{equation}
+	\label{eq:3_17}
+	x^{k+j} =\frac{x_{t_{j} } +x_{t_{j} -1} }{2} -\frac{z_{t_{j} } -z_{t_{j} -1} }{2m}.
+	\end{equation}
+
+	In order to satisfy the assertion of the theorem, let us assume that the minimized function $\varphi(x)$ satisfies the Lipschitz condition with the constant $L>0$, i.e.,
+
+	\begin{equation}
+	\label{eq:3_18}
+	\left|\varphi (x')-\varphi (x'')\right|\le L\left|x'-x''\right|,x',x''\in [a,b],
+	\end{equation}	
+	and, in addition, the method parameter $m>L$.
+
+	Being Lipschitzian, the function $\varphi(x)$ is continuous.
+Hence, the characteristic of the interval will tend to the value $-\varphi (\bar{x})$ when shrinking the interval $(x_{i-1} ,x_{i} )$ to the point $\bar{x}$, i.e., for the feasibility of  (\ref{eq:3_7}) one can take $\mu= 1$ and $c=0$.
+Now let us check if inequality  (\ref{eq:3_9}) holds with given $\mu$ and $c$ for the interval $(x_{i-1} ,x_{i} )$ satisfying  (\ref{eq:3_8}).
+We can use a simple equality
+
+	\begin{equation}
+	\label{eq:3_19}
+	\min \{ z_{i-1} ,z_{i} \} =\frac{1}{2} (z_{i-1} +z_{i} -\left|z_{i-1} -z_{i} \right|)
+	\end{equation}
+	to estimate the method characteristic  (\ref{eq:3_16}), along with the Lipschitz condition and the assumption  (\ref{eq:3_8}) that the length of the interval remains constant and positive, starting from some search iteration:
+	\begin{displaymath}
+	R(i)=m\frac{x_{i} -x_{i-1} }{2} -\frac{z_{i} +z_{i-1} }{2} >L\frac{x_{i} -x_{i-1} }{2} -\frac{z_{i} +z_{i-1} }{2} \ge \frac{\left|z_{i-1} -z_{i} \right|-(z_{i-1} +z_{i} )}{2} =
+	\end{displaymath}
+	\begin{displaymath}
+	=-\min \{ z_{i-1} ,z_{i} \}. 
+	\end{displaymath}
+
+	The obtained inequalities establish the truth of  (\ref{eq:3_9}).
+
+	In order to determine the factor $\nu $ in  (\ref{eq:3_10}) let us estimate the value
+	\begin{displaymath}
+	x_{t_{j} } -x^{k+j} =\frac{x_{t_{j} } -x_{t_{j} -1} }{2} +\frac{z_{t_{j} } -z_{t_{j} -1} }{2m} \le \frac{x_{t_{j} } -x_{t_{j} -1} }{2} +L\frac{x_{t_{j} } -x_{t_{j} -1} }{2m} = 
+	\end{displaymath}
+		\begin{displaymath}
+	 =\left(1+\frac{L}{m} \right)\frac{x_{t_{j} } -x_{t_{j} -1} }{2} . 
+	\end{displaymath}
+
+	An analogous estimate takes place for the interval $(x_{t_{j} -1} ,x^{k+j} )$ as well, therefore, one can take $\nu =0,5(1+L/m)$. Obviously, $\nu >0$ and, because $m>L$, $\nu <1$. \\
+
+	\textbf{Core information Algorithm of Global Search (AGS)} \\
+
+	The decision rule  (\ref{eq:3_2})- (\ref{eq:3_6}) of AGS includes characteristics
+
+	\begin{equation}
+	\label{eq:3_20}
+	R(i)=m(x_{i} -x_{i-1} )+\frac{(z_{i} -z_{i-1} )^{2} }{m(x_{i} -x_{i-1} )} -2(z_{i} +z_{i-1} ),\; 1\le i\le \tau ,
+	\end{equation}
+	and points of new trials calculated according to  (\ref{eq:3_17}). The factor $m$ is determined by the expressions
+
+	\begin{equation}
+	\label{eq:3_21}
+	m=\left\{\begin{array}{c} {rM,M>0} \\ {1,M=0} \end{array}\right. ,
+	\end{equation}
+	\begin{equation}
+	\label{eq:3_22}
+	M=\mathop{\max }\limits_{1\le i\le \tau } \frac{\left|z_{i} -z_{i-1} \right|}{x_{i} -x_{i-1} },
+	\end{equation}
+	where $r>1$ is a parameter of the method.
+
+	Having assumed the Lipschitz property  (\ref{eq:3_18}) of the objective function  (\ref{eq:3_1}), let us prove that this method also satisfies the conditions of Theorem \ref{theor:3_1} with $\mu=4$, $c=0$, and $\nu =0,5(1+1/r)$.
+
+	Because of Lipschitzness when shrinking the interval $(x_{i-1} ,x_{i} )$ into a point $\bar{x}$ its characteristic tends to $-4\varphi (\bar{x})$, i.e., for the truth of  (\ref{eq:3_7}) one can take $\mu=4$ and $c=0$.
+
+	In order to check  (\ref{eq:3_9}) under the condition  (\ref{eq:3_8}) let us consider two cases.
+First, assume that the equality $z_{i-1} =z_{i} $ holds for the interval $(x_{i-1} ,x_{i} )$.
+Then, the characteristic
+	\begin{equation*}
+	R(i)=m(x_{i} -x_{i-1} )-2(z_{i} +z_{i-1} )>\; -2(z_{i} +z_{i-1} )= -4\min \{ z_{i-1} ,z_{i} \},
+	\end{equation*}
+	since the length of the interval $(x_{i-1} ,x_{i} )$ does not change anymore, starting from some iteration $n_{\Delta } $ , i.e., a constant $\Delta >0$ exists such that for all $n>n_{\Delta } $ the inequalities $x_{i} -x_{i-1} >\Delta >0$ hold.
+
+	Now assume that $z_{i-1} \ne z_{i} $. Let us rewrite the characteristic  (\ref{eq:3_20}) in the form
+
+	\begin{equation}
+	\label{eq:3_23}
+	R(i) = |z_{i} -z_{i-1} |(\beta +\frac{1}{\beta } )-2(z_{i} +z_{i-1} ),
+	\end{equation}
+	where $0<\beta =\frac{m(x_{i} -x_{i-1} )}{\; \left|z_{i} -z_{i-1} \right|} \; <1$ as a consequence of  (\ref{eq:3_21}),  (\ref{eq:3_22}).
+In this case, the value $\beta +\frac{1}{\beta } >2$, therefore from  (\ref{eq:3_23})
+
+	\begin{equation*}
+	R(i)>\; 2\left|z_{i} -z_{i-1} \right|-2(z_{i} +z_{i-1} )=-4\min \{ z_{i} ,z_{i-1} \} .
+	\end{equation*}
+
+	As for the condition  (\ref{eq:3_10}), according to  (\ref{eq:3_22}) the following estimate holds: 
+	\begin{displaymath}
+	x_{t_{j} } -x^{k+j} =\frac{x_{t_{j} } -x_{t_{j} -1} }{2} +\frac{z_{t_{j} } -z_{t_{j} -1} }{2m} \le \frac{x_{t_{j} } -x_{t_{j} -1} }{2} +M\frac{x_{t_{j} } -x_{t_{j} -1} }{2m} =
+	\end{displaymath}
+	\begin{displaymath}
+	=\left(1+\frac{1}{r} \right)\frac{x_{t_{j} } -x_{t_{j} -1} }{2}. 
+	\end{displaymath}
+
+	The length of the interval $(x_{t_{j} -1} ,x^{k+j} )$ can be estimated in the same way, therefore, one can take $\nu$ = $0.5(1+1/r)$. Since $r>1$, the condition $0<\nu <1$ is satisfied obviously. \\
+
+	\textbf{Information Algorithm of Global Search with trials at internal points (AGSIP)} \\
+
+	In the framework of this method all trials are executed at internal points of the search domain $[a,b]$.
+The set $\Lambda _{n} $ includes the trial points and the end points $a$ and $b$, from where $\tau =k+1$.
+Morover, $z_{j} =\varphi (x_{j} )$ for $1\le j\le k$ only since at the ends points of the domain the trials are not executed.
+
+	The characteristics $R(i),\; 1\le i\le \tau ,$ of AGSIP are computed in accordance with the expression
+
+	\begin{equation}
+	\label{eq:3_24}
+	R(i)=
+	\left\{
+	\begin{array}{l} 
+	{m(x_{i} -x_{i-1} )+\frac{r(z_{i} -z_{i-1} )^{2} }{M(x_{i} -x_{i-1} )} -2(z_{i} +z_{i-1} ),\; 1<i<\tau ,} \\ 
+	{m(x_{i} -x_{i-1} )-4z_{1} ,\; i=1,} \\ 
+	{m(x_{i} -x_{i-1} )-4z_{k} ,\; i=\tau ,} 
+	\end{array}
+	\right., 
+	\end{equation}
+
+	and the next trial points are computed as
+
+	\begin{equation}
+	\label{eq:3_25}
+	x^{k+j} =\; 0.5(x_{t_{j} } +x_{t_{j} -1} )-
+	\left\{
+	\begin{array}{l} 
+		{0,\; \; t_{j} =1\; \; or \; \; t_{j} =\tau ,} \\ 
+		{(z_{t_{j} } -z_{t_{j} -1} )/(2m),\; 2\le t_{j} \le \tau -1.} 
+	\end{array}
+	\right. 
+	\end{equation}
+
+	Under Lipschitz condition  (\ref{eq:3_18}), we obtain that when shrinking the interval $(x_{i-1} ,x_{i} )$ to a point $\bar{x}$, its characteristic tends to $-4\varphi (\bar{x})$, i.e., for the fulfillment of  (\ref{eq:3_7}), it is possible to take $\mu = 4$ and $c = 0$.
+
+	For the characteristics of the «boundary» intervals $(x_{0} =a,x_{1} )$ and $(x_{\tau -1} ,x_{\tau } =b)$ we have $\mathop{\lim }\limits_{n\to \infty } R(1)>-4z_{1} $ for the left interval and $\mathop{\lim }\limits_{n\to \infty } R(\tau )>-4z_k$ for the right one.
+Since the length of the boundary intervals under the condition  (\ref{eq:3_8}) is always positive, the feasibility of these inequalities and, consequently  (\ref{eq:3_9}) is obvious also.
+
+	Let us consider the characteristic of «internal» interval $(x_{i-1} ,x_{i} ),\; 1<i\le k=\tau -1$. Let us write it in the form 
+	\begin{equation*}
+	R(i)=\; m\Delta _{i} (1-r^{-2} )+\frac{r(\left|z_{i} -z_{i-1} \right|-M\Delta _{i} r^{-1} )^{2} }{M\Delta _{i} } -4\tilde{\xi }_{i},
+	\end{equation*}
+
+	where $\Delta _{i} =x_{i} -x_{i-1} ,\; \tilde{\xi }_{i} =\min \{ z_{i-1,} z_{i} \} $.
+Since the second summand is non-negative, the parameter $r>1$, and $\Delta _{i} >0$ the strict inequality $R(i)>-4\tilde{\xi }_{i} $ takes place, i.e.,  (\ref{eq:3_9}) holds.
+
+	As for  (\ref{eq:3_10}), it is obvious that  $\nu =0.5(1+1/r)$.
+
+	\noindent 
+	\subsection{Everywhere dense convergence}
+
+	\begin{theorem} If $\mu = 0$ in the assertions of Theorem \ref{theor:3_1}, then any point of the search domain is a limit point of the search trial sequence.
+	\label{theor:3_2}
+	\end{theorem}
+	\begin{proof} Let us assume that some point $x'\in [a,b]$ is not a limit point of the search trials sequence generated by the algorithm.
+This means, that starting from some iteration $u\ge 1$, the trials do not fall into the interval $(x_{q-1} ,x_{q} )$, $q=q(n)$, such that $x_{q-1} \le x'\le x_{q} $ anymore.
+If $x'\in (a,b),$ and for some $m$ the point $x'=x^{m} $, two such intervals exist, and any of them can be considered.
+
+	Then, for any iteration $n\ge u$, the inequality
+
+	\begin{equation}
+	\label{eq:3_26}
+	\mathop{\lim }\limits_{n\to \infty } R(q)>c.
+	\end{equation}
+	takes place according to  (\ref{eq:3_9}).
+
+	However, the trial sequence being located inside the interval $[a,b]$, includes at least one converging subsequence.
+For the interval $(x_{s-1} ,x_{s} )$, $s=s(n),$ containing a limit point of this subsequence because of the bilateral convergence the relation  (\ref{eq:3_7}) holds, i.e.,
+
+	\begin{equation}
+	\label{eq:3_27}
+	\mathop{\lim }\limits_{n\to \infty } R(s)=c.
+	\end{equation}
+
+	This means that as a consequence of  (\ref{eq:3_26}),  (\ref{eq:3_27}) a number $w\ge u$ exists, such that for all $n\ge w$ the inequality
+
+	\begin{equation*}
+	R(q)>R(s),
+	\end{equation*}
+
+	takes place that contradicts to the initial hypothesis of the impossibility of falling trials into the interval $(x_{q-1} ,x_{q} )$when $n\ge w\ge u$, according to the rules  (\ref{eq:3_4}),  (\ref{eq:3_5}) of the basic scheme of the parallel characteristical algorithm.
+
+	This contradiction brings to a close the theorem proof.
+	\end{proof}
+
+	Theorem \ref{theor:3_2} establishes the conditions for so-called «everywhere dense» convergence, when the method converges to all points of the search domain including the global minimum points certainly.
+This types of convergence is adequate to such classes of problems, for which computing the estimates of the extremum from a finite number of trials is impossible (for example, for the continuous functions).
+In this case, the convergence to global optimum solution can be provided by the everywhere dense convergence property only.
+
+	Among the examples considered above, scanning method and Kushner's method have such type of convergence.
+Among the other known algorithms, these are the methods described in \cite{3_Gorodetsky, 3_Zilinskas1975}.
+
+		The conditions of the everywhere dense convergence are the sufficient conditions of the convergence to the global optimum of the objective function.
+However, the character of convergence of this kind requires an additional analysis for investigations of the efficiency of the search sequence point distribution.
+One of such approaches is based on getting the estimates of the relative density of the trial distribution in different subdomains of the search domain as compared to the trial point density in a neighborhood of the global minimum \cite{3_Gorodetsky, 3_GoroGri}.
+
+	The theorem below establishes another type of behavior of the parallel characteristical algorithms:
+
+	\subsection{Locally optimal convergence}
+
+	\begin{theorem}
+	\label{theor:3_3}
+		Let us assume that:
+
+		\noindent i)~ the objective function $\varphi (x),x\in [a\, ,\, b],$ of the problem  (\ref{eq:3_1}) is a continuous one;
+
+		\noindent ii)~ $x^{*} $ is a limit point of the search trial sequence generated by a parallel characteristical algorithm in the course of solving the problem  (\ref{eq:3_1});
+
+		\noindent iii) the assertions of Theorem \ref{theor:3_1} hold for the algorithm with $\mu >0$;
+
+		\noindent iv)~ the number $p=p(n)$ of parallel trials at iterations is uniformly bounded, i.e.,
+		\[p(n)\le P\] 
+		where $P\ge 1$ is a predefined integer constant.
+
+		\noindent Then:
+
+		\noindent i)~ $\varphi (x^{k} )\ge \varphi (x^{*} ),k\ge 1;$
+
+		\noindent ii)~if another limit point $x^{**} \ne x^{*} $ exists, then $\varphi (x^{**} )=\varphi (x^{*} );$
+
+		\noindent iii) if the number of local minima of $\varphi (x)$ in the interval $[a,b]$ is finite, then $x^{*} $ is a local minimizer of $\varphi (x)$.
+	\end{theorem}
+
+	\textbf{Proof}.
+Let us assume that $j=j(n)$ is the number of the interval $(x_{j-1} ,\, x_{j} )$ such that $x_{j-1} \le x^{*} \, \le x_{j} $.
+Note that for the characteristic of this interval the relation  (\ref{eq:3_13}) holds as a consequence of Theorem \ref{theor:3_1}.
+
+	Let us prove the first statement. Assume that at the iteration with the number $u\ge 1$ a trial has been executed at the point $x^{q} \in T(u)$ such that
+
+	\begin{equation}
+	\label{eq:3_28}
+	\varphi (x^{q} )<\varphi (x^{*} ).
+	\end{equation}
+	Let us denote as $s=s(n),\; n\ge u$ the number of the point $x^{q} $ in the series  (\ref{eq:3_2}), i.e., $x^{q} =x_{s} $.
+
+	Let us demonstrate that the point $x^{q} $ is also a limit point of the trial sequence.
+If it is not so, since a moment when the trials do not fall into the interval $(x_{s-1} ,\, x_{s} )$ containing this point and for the characteristic $R(s)$ of this interval because of  (\ref{eq:3_9}) we have
+
+	\begin{equation}
+	\label{eq:3_29}
+	\mathop{\lim }\limits_{n\to \infty } R(s)>-\mu \varphi (x^{q} )+c.
+	\end{equation}
+	Note, that if $s=0$, an inequality, analogous to  (\ref{eq:3_29}) takes place for the characteristic $R(s+1)$ of the interval $(x_{s} ,\, x_{s+1} )$.
+
+	Now, taking into account  (\ref{eq:3_13}) and  (\ref{eq:3_28}), one can claim that starting from some iteration, the inequality
+
+	\begin{equation*}
+	R(s)>R(j),
+	\end{equation*}
+	holds that because of decision rule  (\ref{eq:3_4}) contradicts to the impossibility of  executing trials in the interval $(x_{s-1} ,\, x_s)$, i.e., to our assumption that the point $x^q $ is not a limit one of the search trial sequence.
+
+	Since the function $\varphi (x)$ is  continuous, the inequality $\varphi (x)<\varphi (x^{*} )$ is fulfilled in a vicinity $\Theta \subseteq [a,b]$ of the point $x^q$.
+Since the point $x^{q} $ is a limit one, this vicinity will include more than $P+1$ points of the trial sequence $\{ x^k \} $.
+However, in this case more than $P$ intervals formed by these points will have larger characteristics than $R(j)$ from  (\ref{eq:3_13}).
+This means that according to rule ( \ref{eq:3_4}) the trials can not fall into the interval $(x_{j-1} , x_j )$ starting since a moment.
+However, this is impossible since the point $x^{*} $ is a limit one.
+
+	The second assertion of the theorem is a simple consequence of the first one since having assumed, for example, that $\varphi (x^{**} )<\varphi (x^{*} )$ we will get at some iteration a trial $x^{q} $ such that $\varphi (x^q )<\varphi (x^{*} )$.
+
+	The theorem is concluded by proving the point $x^{*} $ to be a local optimum one.
+Let us assume the opposite, namely, that $x^{*} $ is not a local minimum point.
+Then, there exists a nonempty vicinity $\Theta \subseteq [a,b]$ of the point $x^{*} $, in which the function $\varphi (x)$ is strictly monotonous, i.e., the inequality $\varphi (x)<\varphi (x^{*} ),\; x\in \Theta $ takes place either to the left or to the right of $x^{*} $.
+Since the convergence to the point $x^{*} $ is bilateral a point $x^j \in \Theta $ will appear in the trial sequence necessarily such that $\varphi (x^j)<\varphi (x^{*} )$.
+In the case when $x^{*} =a$ or $x^{*} =b$ a unilateral convergence is sufficient for the last inequality to be true.
+This fact contradicts to the first statement and completes the proof of the theorem.
+
+	Thus, if in conditions  (\ref{eq:3_7}),  (\ref{eq:3_9}) the constant $\mu$ is positive (this means that the method \textit{accounts for} the information on the function in the asymptotic), then the behavior of the algorithm becomes more task-oriented, i.e., it seeks for such points only, which have the property of local minimum.
+Moreover, the second statement of the theorem claims that the method cannot converge to the local minima of different depth.
+
+	Among the considered algorithms, the method of broken lines, AGS and AGSIP possess this property.
+Among the other known both sequential and parallel methods, these are the information algorithms \cite{3_GrishaginStrongin_EnginCybernetics, 3_SergeyevLocTun, 3_SergMukhKvasLera, 3_StrMonRus, 3_StrMarkin, 3_StrSergMon2000} as well as the methods described in \cite{3_Pinter, 3_SergGri1994, 3_SergeyevGrishaginOMS}.
+
+	\subsection{Sufficient conditions of global convergence}
+
+	Theorem \ref{theor:3_3} does not guarantee the convergence to the global minimum of the investigated optimization problem.
+In order to get these guarantees (or the sufficient conditions of convergence to global minimum) it is necessary to have additional properties of the objective function.
+With this purpose, let us consider a class of functions $\varphi (x),x\in [a\, ,\, b],$ satisfying the H\"{o}lder condition
+
+	\begin{equation}
+	\label{eq:3_30}
+	|\varphi (x')-\varphi (x'')|\le H\sqrt[{N}]{|x'-x''|} ,x',x''\in [a,b],
+	\end{equation}
+	where $N\ge 1$ is an integer number, and $H>0$ is called the H\"{o}lder constant.
+The functions of the type  (\ref{eq:3_30}) arise, in particular, in the reduction of the multidimensional functions onto an interval using Peano mapping [    ] that allows reducing a multidimensional optimization problem to solving a one-dimensional problem  (\ref{eq:3_1}).
+Note that for \textit{N~}$=$~1 condition  (\ref{eq:3_30}) turns into Lipschitz condition  (\ref{eq:3_18}) with $H=L$.
+
+	\begin{theorem}
+	\label{theor:3_4}
+		Let us assume that a function $\varphi (x),x\in [a, b],$ satisfies the H\"{o}lder condition  (\ref{eq:3_30}).
+Then any global minimizer $x^{*} $ of the function $\varphi (x)$ over $[a, b]$ is a limit point of the search trial sequence generated by the parallel characteristic algorithm while solving the problem  (\ref{eq:3_1}) if the conditions  (\ref{eq:3_7}) and  (\ref{eq:3_10}) of Theorem \ref{theor:3_1} are satisfied and the interval $(x_{i-1} ,x_{i} )$ which has no trials inside, i.e., satisfying   (\ref{eq:3_8}), meets the inequality
+
+	\begin{equation}
+	\label{eq:3_31}
+	\mathop{\lim }\limits_{n\to \infty } R(i)>\frac{\alpha \mu H}{\sqrt[{N}]{2} } \sqrt[{N}]{(x_{i} -x_{i-1} )} -\mu \omega _{i} +c,
+	\end{equation}
+	where the constants $\mu, c$, and $\nu$ satisfy the conditions of Theorem \ref{theor:3_1}, the parameters $\alpha =1$,  if $x_{i-1} \in \{ x^s \} $ and $ x_i \in \{ x^s \} $, or $\alpha =2$ when either $x_{i-1} $ or $x_i $ does not belong to the trial sequence $\{ x^s \} $.
+The value  $\omega _i $ is determined as
+	\begin{equation} 
+	\label{eq:3_32} 
+	\omega _{i} =\left\{
+	\begin{array}{c}
+	{\frac{1}{2} (z_{i-1} + z_i ),\; if \; x_{i-1} \in \{ x^s \} \; and \; x_i \in \{ x^s \} ;} \\ 
+	{ \begin{array}{l} 
+		{z_{i-1} \; ,\; \quad \quad \quad if \; x_{i-1} \in \{ x^s \} \; and \; x_i \notin \{ x^s \} ;} \\ 
+		{z_i \quad \; , \quad \quad \quad if\; x_{i-1} \notin \{ x^s \} \, \; and\; \, x_i \in \{ x^s \} .}
+	  \end{array}}
+ 	\end{array}\right. \; \,  
+	\end{equation} 
+	\end{theorem}
+
+	\textbf{Proof}.
+Let us assume that the global minimum $x^{*} $ is not the limit point of the trial sequence and that $i=i(n)$ is the number of the interval $(x_{i-1} , x_i)$ including this point at the $n$-th search iteration, i.e., $x_{i-1} \le x^{*} \le \, x_i $.
+Then the trials will not fall into this interval since an iteration.
+Being in the interval $(x_{i-1} ,\, x_i )$, the point $x^{*} $ can be expressed in the form $x^{*} =\lambda x_{i-1} +(1-\lambda )x_i , 0\le \lambda \le 1$.
+As a consequence of  (\ref{eq:3_30})
+
+	\begin{equation*}
+	\varphi (x_{i-1} )-\varphi (x^{*} )\le H\sqrt[{N}]{(x^{*} -x_{i-1} )} =H\sqrt[{N}]{(1-\lambda )(x_{i} -x_{i-1} )} ,
+	\end{equation*}
+	\begin{equation}
+	\label{eq:3_33}
+	\varphi (x_{i} )-\varphi (x^{*} )\le H\sqrt[{N}]{(x_{i} -x^{*} )} =H\sqrt[{N}]{\lambda (x_{i} -x_{i-1} )} ,
+	\end{equation}
+
+	hence,
+
+	\begin{equation*}
+	\varphi (x_i)+\varphi (x_{i-1} )\le 2\varphi (x^{*} )+H(\sqrt[{N}]{\lambda } +\sqrt[{N}]{1-\lambda } )\sqrt[{N}]{(x_{i} -x_{i-1} )\le } 
+	\end{equation*}
+	\begin{equation}
+	\label{eq:3_34}
+	\le 2\varphi (x^{*} )+H\sqrt[{N}]{(x_i -x_{i-1} )} \mathop{\max }\limits_{0\le \lambda \le 1} \{ \sqrt[{N}]{\lambda } +\sqrt[{N}]{1-\lambda } \} =2\varphi (x^{*} )+\frac{2H}{\sqrt[{N}]{2} } \sqrt[{N}]{(x_{i} -x_{i-1} )} ,
+	\end{equation}
+	because the function $\sqrt[{N}]{\lambda } +\sqrt[{N}]{1-\lambda } $ reaches its maximum in the interval $[0,1]$ at the point $\lambda =\frac{1}{2} $.
+
+	In the case, when the parameter $\alpha =1$ and $\omega _i =\frac{1}{2} (z_{i-1} +z_i )$, we have
+
+	\begin{equation}
+	\label{eq:3_35}
+	\mathop{\lim }\limits_{n\to \infty } R(i)>(\frac{\mu H}{\sqrt[{N}]{2} } \sqrt[{N}]{(x_i -x_{i-1} )} -\frac{\mu }{2} (z_{i-1} +z_i))+c.
+	\end{equation}
+
+	The last inequality along with  (\ref{eq:3_34}) allows obtaining the estimate
+
+	\begin{equation}
+	\label{eq:3_36}
+	\mathop{\lim }\limits_{n\to \infty } R(i)>-\mu \, \varphi (x^{*} )+c.
+	\end{equation}
+
+	Now assume that $\omega _i =z_i $. Then $\alpha =2$ (the case of $\omega _i =z_{i-1} $ can be considered by analogy). From  (\ref{eq:3_33})
+	\[-\mu z_i \ge -2\mu \varphi (x^{*} )-\frac{2\mu H}{\sqrt[{N}]{2} } \sqrt[{N}]{(x_{i} -x_{i-1} )} +\mu \varphi (x_{i-1} )\] 
+	and, moreover, $\varphi (x_{i-1} )\ge \varphi (x^{*} )$as $x^{*} $ is a global minimize. Therefore,
+
+	\begin{equation*}
+	\mathop{\lim }\limits_{n\to \infty } R(i)>\frac{2\mu H}{\sqrt[{N}]{2} } \sqrt[{N}]{(x_{i} -x_{i-1} )} -\frac{\mu }{2} z_{i} +c\ge 
+	\end{equation*}
+	\begin{equation*}
+	\ge \frac{2\mu H}{\sqrt[{N}]{2} } \sqrt[{N}]{(x_i -x_{i-1} )} -2\mu \varphi (x^{*} )+\mu \varphi (x_{i-1} )-\frac{2\mu H}{\sqrt[{N}]{2} } \sqrt[{N}]{(x_i -x_{i-1} )} +c\ge -\mu \varphi (x^{*} )+c,
+	\end{equation*}
+	i.e.,  (\ref{eq:3_36}) is true again.
+
+	At the same time, the trial sequence has at least one limit point $\bar{x}$ in the interval $[a,b]$, for which the inequality $\varphi (\bar{x})\ge \varphi (x^{*} )$ is true necessarily.
+Besides, according to Theorem \ref{theor:3_1} for the interval $(x_{j-1} ,\, x_j )$, $j=j(n)$ including $\bar{x}$, the equality
+	\begin{equation*}
+	\mathop{\lim }\limits_{n\to \infty } R(j)=-\mu \, \varphi (\bar{x})+c.
+	\end{equation*}
+	takes place.
+However, in this case, the inequality $R(i)>R(j)$ holds starting from some search iteration.
+But because of the rule  (\ref{eq:3_4}) this inequality contradicts to the assumption of the impossibility to place the trials in the interval $(x_{i-1} ,\, x_i)$.
+
+	\begin{corollary} If $\mu =0$,  (\ref{eq:3_31}) coinsides with  (\ref{eq:3_9}) and Theorem \ref{theor:3_4} becomes identical to Theorem \ref{theor:3_1}.
+In this case, H\"{o}lder property of the objective function is not required.
+	\end{corollary}
+	\begin{corollary} If in the conditions of the theorem $\mu >0$ and the number of processors involved in the execution of given iteration is limited uniformly, then the characteristical algorithm converges to all the global minimum points and only to these ones.
+	\end{corollary}
+	\begin{proof} Indeed, the convergence to all global minimum points is a result of the theorem itself.
+The impossibility of the convergence to the points, being different from the global minimizers is a consequence of the second statement of Theorem \ref{theor:3_3}.
+	\end{proof}
+
+	Let us establish the sufficient conditions of convergence for the method of broken lines and AGS, in which the factor $\mu >0$ under assumption  (\ref{eq:3_18}) for the objective function.
+As $N=1$ and $\omega _i =\frac{1}{2} (z_{i-1} +z_{i} )$, condition  (\ref{eq:3_31}) turns into
+
+	\begin{equation*}
+	\mathop{\lim }\limits_{n\to \infty } R(i)>\frac{\mu L}{2} (x_{i} -x_{i-1} )-\mu \omega _i.
+	\end{equation*}
+
+	For the method of broken lines inequality  (\ref{eq:3_31}) is satisfied obviously when $m>L$ if one keeps in mind the relation  (\ref{eq:3_19}).
+
+	For the characteristic of AGS an obvious inequality holds
+
+	\begin{equation*}
+	R(i)\ge m(x_{i} -x_{i-1} )-2(z_{i} +z_{i-1} )=m(x_{i} -x_{i-1} )-\frac{\mu }{2} (z_{i} +z_{i-1} ),
+	\end{equation*}
+	from where the feasibility of  (\ref{eq:3_31}) follows for $m>2L$.
+
+	Concerning the method AGSIP  (\ref{eq:3_24})- (\ref{eq:3_25}), in the case $1<i<\tau $ its characteristic satisfies the previous inequality as well, i.e., $m>2L$.
+However, if $i=1$ then  (\ref{eq:3_31}) is simplified to
+	\[\mathop{\lim }\limits_{n\to \infty } R(1)>4L(x_{1} -x_{0} )-4z_{1} \] 
+	because of $\alpha =2,\mu =4$, and  (\ref{eq:3_31}) is met if $m>4L$. The case $i=\tau $ can be considered analogously. Thus, the global convergence of AGSIP is provided by the inequality $m>4L$.
+
+	\noindent 
+	\section{Conditions of the non-redundant parallelization}
+
+	The parallelization of the sequential characteristical algorithms according to the scheme  (\ref{eq:3_2})--(\ref{eq:3_6}) leads to the loss of a part of search information when computing the coordinates of the next iteration $x^{k+j} $, $1\le j\le p,$ see  (\ref{eq:3_5}).
+As a matter of fact, in a sequential algorithm, to the moment of selecting the (single) point of the next trial, the results of \textit{all} preceding trials (the function values) are known already.
+But in a parallel method, the selection of the points $x^{k+j} $, $1<j\le p,$ is performed without taking into account the function values at the points $x^{k+s} $, $1\le s<j$ since these values have not been computed yet.
+The higher the parallelization level $p(n)$ in scheme  (\ref{eq:3_2})--(\ref{eq:3_5}), the greater the losses of the information while planning the trials.
+These losses are the most essential in the full parallelization, when $p(n)=\tau (n)$ in  (\ref{eq:3_2}).
+
+	An incomplete accounting for the information can result in a situation when a parallel algorithm distributes the trials more dense as compared to the sequential prototype, i.e., it generates \textit{redundant} trials.
+
+	Let us evaluate the redundancy of the parallelization quantitatively following the approach from [66].
+Let us assume that $\{ x^{k} \} $ and $\{ y^{s} \} $ are the infinite trial sequences ($\varepsilon =0$ in the termination conditions  (\ref{eq:3_14}) generated by a pure sequential characteristic method and by its parallel generalization, respectively, when these two solve the same problems  (\ref{eq:3_1}).
+A coincidence of these sequences, i.e.,
+	\begin{equation} 
+	\label{eq:3_37} 
+	\{ x^{k} \} =\{ y^{s} \}  
+	\end{equation} 
+	means that the parallel algorithm distributes the trials at the same points as the pure sequential one does.
+Note that  (\ref{eq:3_37}) does not require  step-by-step coincidence $x^q =y^q ,\; q\ge 1$.
+
+	When relation  (\ref{eq:3_37}) takes place, the parallelization is called a \textit{non-redundant} one.
+If  (\ref{eq:3_37}) does not hold and $\{ x^{k} \} \subseteq \{ y^{s} \} $, then the parallel scheme has some redundancy.
+Let us introduce a \textit{redundancy coefficient} in order to characterize the redundancy quantitatively as
+	\begin{equation}
+	\label{eq:3_38}
+	\lambda (s,q)=V(s,q)/(s-q),\; s>q,
+	\end{equation}
+	where
+	\begin{equation*}
+	V(s,q)=card(\{ y^{q+1} ,\ldots ,y^{s} \} \backslash \{ x^{k} \} )
+	\end{equation*}
+
+	is the number of the redundant points in $\{ y^{s} \} $ from the (\textit{q}+1)-th trial to the \textit{s-th} one.
+The definition  (\ref{eq:3_38}) supposes the inclusion $\{ x^k \} \subseteq \{ y^s \} $.
+Obviously, $V(s, 0)=0$ for any $s\ge 1$ corresponds to the non-redundant case  (\ref{eq:3_37}).
+
+	Now let us consider the parallel characteristical algorithms, which after $n{}_{0} \ge 1$ initial  iterations use always two processors, i.e.,
+	\begin{equation}
+	\label{eq:3_39}
+	p(n)=2,\; n>n_{0}.
+	\end{equation}
+
+	Let us make several non-burdensome assumptions.
+Assume that if the trial point $x^{k+q} $ of current iteration is selected in an interval $(x_{j-1} ,x_{j} )$ such that the objective function values $z_{j-1} ,z_j $ have been computed already in the ends of this interval, then this point can be represented in the form
+	\begin{equation}
+	\label{eq:3_40}
+	x^{k+q} =g_{j} +sign(z_{j-1} -z_{j} )\theta _{j} ,
+	\end{equation}
+	where
+	\begin{equation}
+	\label{eq:3_41}
+	g_{j} =(x_{j-1} +x_{j} )/2, 
+	\end{equation}
+	\begin{equation}
+	\label{eq:3_42}
+	0\le \theta _{j} \le \sigma (x_{j} -x_{j-1} ),
+	\end{equation}
+	and $0<\sigma \le 0.5$. Note that  (\ref{eq:3_10}) follows from  (\ref{eq:3_40}) -  (\ref{eq:3_42}) for $\nu =0.5+\sigma $.
+
+	If the objective function value has not been computed at one of the ends of the interval $(x_{j-1} ,x_{j} )$ then 
+	\begin{equation}
+	\label{eq:3_43}
+	x^{k+q} =g_{j} .
+	\end{equation}
+
+	Along with  (\ref{eq:3_40}) -  (\ref{eq:3_43}), assume that the characteristics $R(i), 1\le i\le \tau $, and the values $\theta _j ,\; 1\le j\le p$, depend on the boundary points of the corresponding interval and on the function values computed in these points only, i.e.,
+	\begin{equation}
+	\label{eq:3_44}
+	R(i)=\Sigma (x_{i-1} ,x_{i} ,z_{i-1} ,z_{i} ),
+	\end{equation}
+	\begin{equation}
+	\label{eq:3_45}
+	\theta _{j} =\Theta (x_{j-1} ,x_{j} ,z_{j-1} ,z_{j} ).
+	\end{equation}
+
+	Finally, let us assume that
+
+	\begin{equation}
+	\label{eq:3_46}
+	y^s =x^s ,\; 1\le s\le k(n_{0} ),
+	\end{equation}
+	i.e., the initial trials, preceding the characteristical rule  (\ref{eq:3_2}) - (\ref{eq:3_5}), are the same for both sequential and parallel methods.
+
+	\begin{theorem} 
+	\label{theor:3_5}
+	Let us assume that: \\
+	i) the function $\varphi (x),x\in [a\, ,\, b],$ to satisfy Lipschitz condition  (\ref{eq:3_18}) with the constant  $L>0$; \\
+	ii) the conditions  (\ref{eq:3_39}),  (\ref{eq:3_40}),  (\ref{eq:3_42}) -  (\ref{eq:3_45}), and  (\ref{eq:3_7}) hold for the parallel characteristic algorithm and for its sequential prototype; \\	
+	iii) the constants $\mu $ and $\sigma $ satisfy the inequalities $\mu >0$, $\sigma \le \frac{1}{6} $; \\
+	iv) for any interval $(x_{i-1} ,\, x_i )$ with the length $\Delta _i >0$ its characteristic 
+	\begin{equation}
+	\label{eq:3_47}
+	R(i)>\mu L(x_{i} -x_{i-1} )-\mu \tilde{\xi }_{i} +c.
+	\end{equation}
+
+	Then for the redundancy coefficient  (\ref{eq:3_38}) the estimate 
+	\begin{equation} 
+	\label{eq:3_48} 
+	\lambda (s,n_{0} )\le E[(s-n_{0} )/6]/(s-n_{0} )<0.17 
+	\end{equation} 
+	is true. Here $E[\vartheta ]$ is the integer part of $\vartheta $.
+	\end{theorem}
+	\textbf{Proof.} If after the $s$-th iteration of the parallel method and the $k$-th iteration of the sequential one the equality 
+	\begin{equation}
+	\label{eq:3_49}
+	(y_{t-1} ,y_{t} )=(x_{q-1} ,x_{q} ),\; t=t(s),\; q=q(k),
+	\end{equation}
+	holds and the next trial points $y^{s+1} $ and $x^{k+1} $ fall into the given intervals, respectively, then as a consequence of  (\ref{eq:3_44})
+	\begin{equation}
+	\label{eq:3_50}
+	R(t(s))=R(q(k)),
+	\end{equation}
+	and, according to  (\ref{eq:3_6}),  (\ref{eq:3_40}), and  (\ref{eq:3_42}), 
+	\begin{equation}
+	\label{eq:3_51}
+	y^{s+1} =x^{k+1}.
+	\end{equation}
+
+	The decision rules of the characteristical algorithms along with  (\ref{eq:3_49})- (\ref{eq:3_51}) and  (\ref{eq:3_37}) with $s=k=n_0 $ provide the inclusion
+	\begin{equation*}
+	\{ x^{k} \} \subseteq \{ y^{s} \},
+	\end{equation*}
+	which allows evaluating the redundancy with the help of the coefficient  (\ref{eq:3_38}).
+
+	Conditions  (\ref{eq:3_7}),  (\ref{eq:3_47}), and  (\ref{eq:3_10}) with $\nu =0,5+\sigma $ ensure the fulfillment of the conditions of Theorem \ref{theor:3_4}, according to which when $\mu >0$ the global minimum points of the function $\varphi (x)$ only can be the limit points of the sequences $\{ x^k \} $and\textit{ $\{ y^{s} \} $}.
+This means that the set of limit points of the sequence \textit{$\{ y^{s} \} $} coincides with the set of limit points of the sequence $\{ x^k \} $.
+
+	Assume that the first \textit{k} points of the sequence $\{ x^k \} $ are ordered according to  (\ref{eq:3_3}) and $j=j(k)$ is the number of an interval $(x_{j-1} ,x_j )$ such that $x_{j-1} \le x^{*} \le x_{j} $ where $x^{*} $ is the global minimizer of the function $\varphi (x)$.
+On account of  (\ref{eq:3_47}) and Lipschitz condition, we have the inequality
+	\begin{equation}
+	\label{eq:3_52}
+	R(j)>-\mu \varphi (x^{*} )+c.
+	\end{equation}
+	for the characteristic of this interval. Since the point $x^{*} $ is a limit one, then according to  (\ref{eq:3_7}) 
+	\begin{equation*}
+	R(j(k))\to -\mu \varphi (x^{*} )+c+0.
+	\end{equation*}
+	if $k\to \infty $.
+
+	Taking into account the rule  (\ref{eq:3_4}) one can state that any interval $(x_{i-1} ,x_{i} ),\; i=i(k),$ the characteristic of which satisfies  (\ref{eq:3_52}) includes at least one point of the sequence $\{ x^{k} \} $.
+At the same time, any interval $(x_{i-1} ,x_{i} ),\; i=i(k),$ for which
+	\begin{equation}
+	\label{eq:3_53}
+	R(i)<-\mu \varphi (x^{*} )+c,
+	\end{equation}
+	cannot include the points of the sequence $\{ x^{k} \} $. Thus, the redundant trials of the parallel method can be executed only in the intervals, for which  (\ref{eq:3_53}) holds.
+
+	Let us consider again the interval \textit{$[x_{j-1} ,x_{j} ],\; j=j(k)$}including the global minimum point $x^{*} $.
+The current trial at the point $x=x^{k+1} $ within this interval forms two new subintervals
+	\begin{equation}
+	\label{eq:3_54}
+	[x_{j-1} ,\; x]\, ,\; [x,x_{j} ],
+	\end{equation}
+	one of which includes $x^{*} $. Let it be the first subinterval, i.e.,
+	\begin{equation}
+	\label{eq:3_55}
+	x^{*} \in [x_{j-1} ,\; x].
+	\end{equation}
+
+	Then the inequality  (\ref{eq:3_52}) is true for this interval. Let us show that this inequality holds for the interval \textit{$\; [x,x_{j} ]$} as well. Let us introduce the notations
+	\begin{equation}
+	\label{eq:3_56}
+	\Delta _{x} =x_{j} -x,\; \xi _{x} =\; \min \{ z_{j} ,\; z\} ,\; z=z^{k+1} .
+	\end{equation}
+
+	Let us consider the case when the points $x_{j-1} $ and $x_j $ are the coordinates of the search trials, i.e., the function values $z_{j-1} $ and $z_j $ have been computed in these points, therefore, $\tilde{\xi }_{j} =\min \{ z_{j-1} ,z_j \} $.
+Let us estimate the value of $\xi _x $.
+Assume first that $z_{j-1} <z_j $.
+On account of  (\ref{eq:3_55}),  (\ref{eq:3_40}),  (\ref{eq:3_42}), and  (\ref{eq:3_56}) we have
+	\begin{equation*}
+	x^{*} \le x=g_j -\theta _j \le g_j 
+	\end{equation*}
+	\begin{equation*}
+	\xi _x \le z\le \varphi (x^{*} )+L(x-x^{*} )<\varphi (x^{*} )+L\Delta _x.
+	\end{equation*}
+
+	If $z_{j-1} \ge z_j $, let us consider two cases. In the first case when $x^{*} >g_{j} $ we have 
+	\begin{equation*}
+	x=g_j +\theta _j 
+	\end{equation*}
+	\begin{equation*}\xi _x \le z\le \varphi (x^{*} )+L(x-x^{*} )<\varphi (x^{*} )+L(x-g_j )=
+	\end{equation*}
+	\begin{equation*}
+	=\varphi (x^{*} )+L\theta _{j} \le \varphi (x^{*} )+L\sigma \Delta _{j} \le \varphi (x^{*} )+L\Delta _{x} ,
+	\end{equation*}	
+	since $\Delta _x \ge (1-\nu )\Delta _j =(0,5-\sigma )\Delta _j $ (see Note to Theorem \ref{theor:3_1}), and because of $\sigma \le {1/{\vphantom{1 6}}6} $
+
+	\begin{equation*}
+	\Delta _{x} \ge \frac{1}{3} \Delta _{j} >\sigma \Delta _{j}.
+	\end{equation*}
+
+	In the second case when $x^{*} \le g_j $ we can build a chain of inequalities
+	\begin{equation*}
+	\xi _{x} =0,5(z_{j} +z-\left|z_{j} -z\right|)\le 0,5(z_{j} +z)\le 0,5(z_{j-1} +z)\le \phi (x^{*} )+0,5L(x-x_{j-1} )\le 
+	\end{equation*}
+	\begin{equation*}
+	\le \varphi (x^{*} )+\frac{L(1+2\sigma )}{2(1-2\sigma )} \Delta _{x} \le \varphi (x^{*} )+L\Delta _{x},
+	\end{equation*}
+	which allows estimating the characteristic \textit{$R_{x} $} of the interval $(x,x_{j} )$ as
+	\begin{equation*}
+	R_{x} >\mu L\Delta _{x} -\mu \xi _{x} +c\ge -\mu \varphi (x^{*} )+c,
+	\end{equation*}
+	from where 
+	\begin{equation}
+	\label{eq:3_57}
+	R_{x} >-\mu \varphi (x^{*} )+c.
+	\end{equation}
+
+	The inequality  (\ref{eq:3_57}) holds as well when $x_{j} \notin \{ x^{s} \} $ because in this case $x=g_j $ according to  (\ref{eq:3_43}).
+Since $\varphi (x)$ satisfies the Lipschitz condition, $z\le \varphi (x^{*} )+L\Delta _x $ therefore
+	\begin{equation*}
+	R_{x} >\mu L\Delta _{x} -\mu z+c\ge -\mu \varphi (x^{*} )+c.
+	\end{equation*}
+
+	Thus, if a pair of intervals of type  (\ref{eq:3_54}) exists after k trials then the choice of two trials simultaneously according to the decision rules  (\ref{eq:3_2}) -  (\ref{eq:3_6}) and  (\ref{eq:3_39}) cannot produce redundant trials.
+However, after $k(n_0)$ initial trials  (\ref{eq:3_46}) we can obtain the worst situation when there exists only one interval satisfying the condition  (\ref{eq:3_52}), namely, the interval $[x_{j-1} ,x_j ]$, which includes the global minimizer $x^{*} $ and the other intervals  violate the inequality  (\ref{eq:3_52}).
+This means that the interval \textit{$[x_{j-1} ,x_{j} ]$} has the maximal characteristic among all ones, therefore the first trial of the new $(n_{0} +1)$-th iteration will be executed in this interval according to rule  (\ref{eq:3_4}) but the second trial of this iteration will be  redundant.
+
+	The execution of a trial in the interval $[x_{j-1} ,x_j ]$ will result in an appearance of a pair of intervals  (\ref{eq:3_54}), and  (\ref{eq:3_52}) holds for each of them.
+Therefore, at the $(n_0 +2)$-th iteration the next two trials $y^q \in \{ x^k \} $ will be executed in these intervals.
+If one of the trials of the $(n_0 +3)$-th iteration falls into the interval  (\ref{eq:3_55}), this will result in the appearance of a new pair  (\ref{eq:3_54}) again and in the absence of the redundant trials at this iteration.
+Thus, the source of redundant trials can be  the situation when a pair of trials of the $(n_0 +3)$-th iteration includes the points from \textit{$(x,x_j )$} but does not contain any point from  (\ref{eq:3_55}).
+If this situation takes place at the $(n_0 +3)$-th iteration then the $(n_0 +4)$-th iteration can execute one trial in interval  (\ref{eq:3_55}) and another one in an interval $(x_{i-1} ,x_i )$ such that its characteristic satisfies the inequality  (\ref{eq:3_53}).
+The latter trial will be  redundant.
+After that, we have the situation identical to the one occurred after the execution of the $(n_0 +1)$-th iteration.
+
+	Thus, during three iterations from the $(n_0 +1)$-th to the $(n_0 +3)$-th one inclusive, in which 6 trials have been executed, one can obtain  no more than one redundant trial.
+This means that relation  (\ref{eq:3_48}) is true.
+
+	\begin{corollary} If the objective function $\varphi (x)$ has $\gamma $ global minimum points and there is at least one point of the sequence $\{ y^q \} $ (generated during the execution of the initial iterations) between two adjacent global minimizers, then under the conditions of Theorem \ref{theor:3_5} the parallel characteristical algorithm with $2\gamma $ parallel processors involved into every iteration provides the redundancy according to  (\ref{eq:3_48}).
+	\end{corollary}
+
+	\section{ Experimental results}
+
+	In order to illustrate the theoretical conclusions drawn above the results of the experiments carried out using a multiprocessor cluster at the Lobachevsky State University of Nizhni Novgorod and mini-supercomputer ALLIANT FX/80 on the test classes of functions of various dimensionality are presented in this subsection.
+
+	Let us assume that the computational time costs of a trial are much greater than the ones of the implementation of the algorithm computational scheme so that the latter can be neglected.
+In this situation, one can use the speed up in iterations
+	\begin{equation*}
+	s(p)=\frac{K(1)}{K(p)/p} ,\; p>1
+	\end{equation*}
+
+	as a general measure for estimating the parallelization efficiency.
+Here $K(1)$ is a number of trials executed by the sequential characteristical algorithm during solving an optimization problem (after the completion of the initial trials) and $K(p)$ is the number of trials executed by the parallel method using \textit{p} parallel processors at every iteration during solving the same problem.
+
+	For carrying out the experiments with the one-dimensional parallel characteristical algorithms a set of the test functions proposed in \cite{3_HansenJaumardLu} has been used.
+The parallel algorithms based on the sequential versions of the Kushner's method (KM), of the Method of Broken Lines (MBL), and of the information algorithm AGSIP have been considered.
+On the base of three functions from \cite{3_HansenJaumardLu} and various starting points, three problems have been taken:
+
+	PROBLEM 1. Function \#2, starting points 3, 4, and 5.
+
+	PROBLEM 2. Function \#9, starting points 5, 8, and 11.
+
+	PROBLEM 3. Function \#19, starting points 1, 2, and 3.
+
+	For Kushner's method and for method of broken lines the initial trials have been executed at the boundary points $a$ and $b$ of the corresponding search intervals $[a,b]$ as well.
+For these methods, the exact values of Lipschitz constants have been used, unlike AGSIP, for which the adaptive estimate  (\ref{eq:3_21}) with $r=2$ has been employed.
+The value of $\epsilon = 0.0001(b-a)$ has been used in the termination criterion  (\ref{eq:3_14}) for these methods.
+The results of the experiments are summarized in Table~\ref{tab:3_1} where $q$ denotes the problem number and $p$ is the number of the parallel processors.
+The global minima were found in all cases.
+
+	\begin{table}[ht]
+		\centering
+		\caption{The result of experiments for the invariant functions}
+		\label{tab:3_1}
+		{\setlength{\extrarowheight}{1.5pt}
+			\begin{tabular}
+				{|c|c|c|c|c|c|c|c|}
+				\hline
+				\multirow{2}{*}{\textbf{$q$}} 
+				& \multirow{2}{*}{\textbf{$p$}} 
+			    	 & \multicolumn{2}{|c|}{\textbf{KM}} 
+				     & \multicolumn{2}{|c|}{\textbf{MBL}} & \multicolumn{2}{|c|}{\textbf{AGSIP}} \\
+				\cline{3-8}
+				  &  & \textbf{Number of trials} & \textbf{Speed up} 
+				     & \textbf{Number of trials} & \textbf{Speed up} 
+				     & \textbf{Number of trials} & \textbf{Speed up} \\ 
+				\hline
+				\multirow{4}{*}{1} 
+				  & 1 & 4252 & --   & 151 & --   & 120 & --   \\ \cline{3-8}
+				  & 2 & 4353 & 1.95 & 151 & 2.0  & 119 & 2.02 \\ \cline{3-8}
+				  & 3 & 4352 & 2.93 & 155 & 2.92 & 162 & 2.21 \\ \cline{3-8}
+				  & 4 & 4353 & 3.91 & 157 & 3.83 & 163 & 2.93 \\ \hline
+				\multirow{4}{*}{2} 
+				  & 1 & 4622 & --   & 124 & --   & 125 & --   \\ \cline{3-8}
+				  & 2 & 4621 & 2.0  & 123 & 2.02 & 125 & 2.0  \\ \cline{3-8}
+				  & 3 & 4622 & 3.0  & 122 & 3.05 & 126 & 2.98 \\ \cline{3-8}
+				  & 4 & 4373 & 4.24 & 125 & 3.7  & 127 & 3.94 \\ \hline
+				\multirow{4}{*}{3} 
+				  & 1 & 2780 & --   & 126 & --   & 148 & --   \\ \cline{3-8}
+				  & 2 & 2799 & 2.0  & 127 & 1.98 & 153 & 1.93 \\ \cline{3-8}
+				  & 3 & 2762 & 3.02 & 128 & 2.95 & 159 & 2.79 \\ \cline{3-8}
+				  & 4 & 2769 & 4.02 & 133 & 3.78 & 163 & 3.63 \\ \hline
+
+			\end{tabular}
+		}
+
+	\end{table}
+
+	As follows from the results of the experiments, in some cases the methods have demonstrated greater speed up than the number of processors used (superlinear speed up).
+This situation is possible in the Kushner's method and in AGSIP because the objective function behavior is estimated adaptively in these algorithms.
+For example, in AGSIP the hyperspeed up will take place if Lipschitz constant is estimated more precisely in the parallel version than in the sequential one.
+
+	Further experiments have been carried out for the multidimensional problems according to the approach to the dimensionality reduction based on Peano curves presented in Chapter 5.
+The second series of experiments was devoted to maximization of 100 two-dimensional functions of the type
+	\begin{equation}
+	\label{eq:3_58}
+	\varphi(y)=\left\{\left(\sum _{i=1}^7\sum _{j=1}^7A_{ij} g_{ij} (y)+B_{ij}h_{ij}(y) \right)^2 +\left(\sum _{i=1}^7\sum _{j=1}^7C_{ij}g_{ij}(y)-D_{ij}h_{ij}(y)\right)^2 \right\}^{1/2},
+	\end{equation}
+	where 
+	\begin{equation*}
+	y=(y_1 ,y_2 )\in R^2 ,0\le y_s \le 1,s=1,2,
+	\end{equation*}
+	\begin{equation*}
+	g_{ij} (y)=\sin (i\pi y_1 )\sin (j\pi y_2 ),
+	\end{equation*}
+	\begin{equation*}
+	h_{ij} (y)=\cos (i\pi y_1 )\cos (j\pi y_2 )
+	\end{equation*}
+	 and the coefficients $A_{ij} ,B_{ij} ,C_{ij} ,D_{ij} $ were distributed uniformly over the interval $[-1,1]$.
+
+	The test problems were reduced to the one-dimensional ones by Peano mappings.
+Two parallel characteristic algorithms were applied to solve the reduced one-dimensional problems until the precision $\varepsilon =0.001$  in the termination condition is reached.
+The first algorithm was described in \cite{3_SergGri1994}.
+It uses a fixed estimate of Lipschitz constant (let us denote this algorithm as Parallel Method with Fixed Lipschitz constant, or PMFL).
+Note that obtaining this estimate has required 2000 computations of the function values on a uniform grid.
+Using PMFL, the global  solutions have been found in 99 cases of 100.
+In the only case, the solution has not been found (neither by a sequential algorithm nor by the parallel method) because of not enough accurate estimate of the Lipschitz constant: the estimate obtained was less than the actual value.
+The second algorithm, used for comparison, was the parallel multidimensional information algorithm applying Peano curves (PMIP).
+Using PMIP, the optima have been found in all 100 problems.
+The results of the experiments (the averaged number of trials and the speed up) are presented in Table ~\ref{tab:3_2}.
+
+	\begin{table}[ht]
+		\centering
+	\caption{Averages results for two-dimensional functions  (\ref{eq:3_58})}
+	\label{tab:3_2}
+	{\setlength{\extrarowheight}{1.5pt}
+	\begin{tabular}
+		{|c|c|c|c|c|}
+		\hline
+		\multirow{2}{*}{\textbf{P}} & \multicolumn{2}{|c|}{\textbf{PMFL}}  & \multicolumn{2}{|c|}{\textbf{PMIP}} \\ \cline{2-5}
+		 & \textbf{Number of trials} & \textbf{Speed up} & \textbf{Number of trials} & \textbf{Speed up} \\ \hline
+		1 & 1013.25 & --    & 1575.12 & --    \\ 		\hline
+		2 & 1012.82 & 2.001 & 1569.08 & 1.974 \\ 		\hline
+		3 & 1012.33 & 3.003 & 1562.61 & 3.024 \\ 		\hline
+		4 & 1012.00 & 4.005 & 1599.92 & 3.938 \\ 		\hline
+	\end{tabular}
+}
+	\end{table}
+
+	The third series of experiments consisted in the minimization of the multidimensional function 
+	\begin{equation}
+	\label{eq:3_59}
+	\varphi (y)=\frac{\pi }{N} \left\{\sin ^{2} (\pi z_{1} )+5(z_{N} -1)^2 +\sum _{i=1}^{N-1}\left[(z_{i} -1)^2 (1+\sin ^2 (\pi z_{i+1} ))\right] \right\}
+	\end{equation}
+	where $z_{i} =1+0,25(y_i -1)$, $-10\le y_{i} \le 10$, and $1\le i\le N$ for the dimensions $N=$~2,~3,~and~4. This function is a modification of the test one from \cite{3_LucidiPiccioni}.
+
+	Two methods have been applied for the minimization of this function: PMIP and method of broken lines generalized for the multidimensional case according to the dimensionality reduction scheme based on Peano curves (MBLP).
+The value of $\varepsilon =10^{-5} $ was used in the termination criterion  (\ref{eq:3_14}) of the methods.
+Lipschitz constant was estimated adaptively according to  (\ref{eq:3_21}).
+The parameters $r=3.5$ for PMIP and  $r=1.7$ for MBLP were selected according to the sufficient conditions of convergence for the algorithms.
+The global minima points have been found in all cases.
+The results of the experiments are presented in Table \ref{tab:3_3}.
+
+	\begin{table}[ht]
+		\centering
+		\caption{The results of the comparison for the problem  (\ref{eq:3_59})}
+		\label{tab:3_3}
+		{\setlength{\extrarowheight}{1.5pt}
+		\begin{tabular}
+			{|c|c|c|c|c|c|}
+			\hline
+			\multirow{2}{*}{\textbf{$N$}} 
+			& \multirow{2}{*}{\textbf{$p$}} 
+			& \multicolumn{2}{|c|}{\textbf{PMIP}} 
+			& \multicolumn{2}{|c|}{\textbf{MBLP}} \\
+			\cline{3-6}
+			&  & \textbf{Number of trials} & \textbf{Speed up} 
+			& \textbf{Number of trials} & \textbf{Speed up} \\ 
+			\hline
+			\multirow{4}{*}{2}
+			& 1 & 1082 & --    & 1027 & --    \\ \cline{3-6}
+			& 2 & 1084 & 1.996 & 1028 & 1.998 \\ \cline{3-6}
+			& 3 & 1083 & 2.997 & 1050 & 2.934 \\ \cline{3-6}
+			& 4 & 1136 & 3.810 & 1224 & 3.356 \\ \hline
+			\multirow{4}{*}{3}
+			& 1 & 4784 & --    & 4649 & --    \\ \cline{3-6}
+			& 2 & 4776 & 2.003 & 5018 & 1.853 \\ \cline{3-6}
+			& 3 & 4776 & 3.005 & 4647 & 3.001 \\ \cline{3-6}
+			& 4 & 4768 & 4.016 & 4480 & 4.151 \\ \hline
+			\multirow{4}{*}{4}
+			& 1 & 7958 & --    & 6274 & --    \\ \cline{3-6}
+			& 2 & 7876 & 2.021 & 5238 & 2.396 \\ \cline{3-6}
+			& 3 & 6750 & 3.537 & 6486 & 2.902 \\ \cline{3-6}
+			& 4 & 7820 & 4.072 & 6468 & 3.880 \\ \hline
+		\end{tabular}
+	}
+	\end{table}
+
+	As it has been already mentioned above, the adaptive estimate of the function behaviour during the search process can result in a superlinear speed up.
+In general, the experiments have confirmed the main theroretical conclusions on the convergence of the parallel algorithms and on the efficiency of their parallelizaion drawn in the present chapter.
+
+	\section{Parallel asynchronous characteristical global search algorithms}
+
+	As it has been pointed out above, the synchronous scheme has a disadvantage: the processors completed their trials during current iteration before the others stay idle waiting for final completing the iteration.
+However, an asynchronism can be introduced into the structure of the characteristical methods that allows excluding the idle of processors because of differences in the run time for different processors and, therefore, increasing the parallelization efficiency in comparison with the synchronous scheme.
+
+	Before describing the structure of the asynchronous characteristical algorithm, let us come back to the general computational scheme of the asynchronous method (see (\ref{eq:1_26})--(\ref{eq:1_30})) and remind a number of important aspects and notations.
+Since the iterations are distributed in time in the asynchronous algorithm, it is convenient to use the number of trial \textit{k} when describing the method.
+In this connection, let us introduce a set of the coordinates of the completed trials
+	\begin{equation} 
+	\label{eq:3_60} 
+	w(k)=\{ w_{1} ,w_{2} ,\ldots ,w_{k} \} , 
+	\end{equation} 
+	$k\ge 1$, and a set of the coordinates, at which the new trials are executed after completing the preceding $k$ trials
+	\begin{equation} 
+	\label{eq:3_61} 
+	v(k)=\{ v_{1} ,v_{2} ,\ldots ,v_{p(k)} \}  
+	\end{equation} 
+	Here $p(k)$ is a number of processors executing the current trials, 
+	\begin{equation*}
+	w_{j} =x^{i_{j} } ,\; 1\le j\le k,
+	\end{equation*}
+	\begin{equation*}
+	v_{s} =x^{i_{s} } ,\; 1\le s\le p(k),\; 1\le i_{j} ,\; i_{s} \le k+p(k).
+	\end{equation*}
+
+	It is obvious that
+	\begin{equation*}
+	w(k)\bigcup v(k)=\{ x^{1} ,\ldots ,x^{k} ,\ldots ,x^{k+p(k)} \}.
+	\end{equation*}
+
+	\textbf{Definition 3.2.} An algorithm for solving the problem  (\ref{eq:3_1}) is called  \textit{parallel asynchronous characteristical method} if the decision rule of the trial coordinate selection for the algorithm consists in the following.
+
+	The first $k_{1} \ge 1$ trials are executed in the interval $[a,b]$ arbitrarily, but starting from the search step with number $k=k_{0} \le k_{1} {\kern 1pt} ,\; k_{1} -k_{0} \le p(k_{0} )$, such that $k_{0} $ trials have been completed and $k_{1} -k_{0} $ trials are running, the rule of new trial coordinates selection is as follows:
+
+	\begin{description} [\textbf{Step1.}]
+	\item[\textbf{Step 1.}] {If the idle processors are absent, the algorithm is waiting a free processor.
+If $1\le \pi =\pi (k)\le p(k)$ processors have completed the execution of trials and the function values $z^{i_{j} } =\varphi (x^{i_{j} } )$ at the points $x^{i_{1} } ,\; x^{i_{2} } ,\ldots ,x^{i_{\pi } } $, $x^{i_{j} } \in v(k),\; 1\le j\le \pi ,$ have been computed, then the step number is set to $k=k+\pi $, the sets $w(k)=w(k-\pi )\cup \{ x^{i_{1} } ,\; x^{i_{2} } ,\ldots ,x^{i_{\pi } } \} ,$ $v(k)=v(k-\pi )\backslash \{ x^{i_{1} } ,\; x^{i_{2} } ,\ldots ,x^{i_{\pi } } \} $ are formed, and the transition to the next step is performed.}
+	\item [\textbf{Step 2.}]   {A set 
+	\begin{equation*}
+	\Lambda _{k} =\{ x_{0} ,x_{1} ,...,x_{\tau } \}
+	\end{equation*}
+	of a finite number $\tau +1=\tau (n)+1$ of points in the domain $Q=[a,b]$ is defined provided that $a\in \Lambda _k ,b\in \Lambda _k $, $w(k)\subseteq \Lambda _k ,\; v(k)\subseteq \Lambda _k ,$ and the set $\Lambda _k $ is the ordered one (by the subscript) in the increasing order, i.e.,
+	\begin{equation*}
+	a=x_0 <x_1 <...<x_{\tau -1} <x_{\tau } =b.
+	\end{equation*}
+	}
+	\item [\textbf{Step 3.}] {A list $\Sigma _k$ of the numbers of the feasible intervals $(x_{i-1} ,x_i )$ is formed; such intervals are those, boundary points $x_{i-1} \; {\rm and}\; x_i $ of which are either the coordinates of the completed trials or the boundary points of the interval $[a,\, b]$.
+In other words, the index $i$ of the interval $(x_{i-1} ,x_i )$ belongs to the set $\Sigma _k $ if none of the ends of the interval is a coordinate of a noncompleted trial.
+}
+
+	\item [\textbf{Step 4.}] {The value $R(i)$ is juxtaposed to each feasible interval $(x_{i-1} ,x_i )$, $i\in \Sigma _k $. $R(i)$ is called the characteristic of this interval. }
+
+	\item [\textbf{Step 5.}] {The characteristics $R(i),\; 1\le i\le \tau $, are arranged in the decreasing order:
+	\begin{equation}
+	\label{eq:3_62}
+	R(t_{1} )\ge R(t_{2} )\ge \ldots \ge R(t_{\tau -1} )\ge R(t_{\tau } ).
+	\end{equation}
+	}
+	\item [\textbf{Step 6.}] {$\pi$ maximal characteristics with the indices $t_j ,\; 1\le j\le \pi $, are selected in the sequence  (\ref{eq:3_62}) and the trials are executed at the points of the set 
+	\begin{equation}
+	\label{eq:3_63}
+	T(k)=\{ x^{\eta +1} ,\ldots ,x^{\eta +\pi } \} ,
+	\end{equation}
+	in the intervals corresponding to these characteristics according to the rules 
+	\begin{equation*}
+	x^{\eta +j} =d(t_{j} )\in (x_{t_{j} -1} ,x_{t_{j} } )
+	\end{equation*}	
+	for $1\le j\le \pi $, where $\eta =card(w(k))+card(v(k))$.
+	}
+	\item [\textbf{Step 7.}] {The points of set $T(k)$ from  (\ref{eq:3_63}) are added to the set \textit{v}(\textit{k}) and the transition to Step 1 is performed.}
+	\end{description}
+
+	The theoretical results connected with the asynchronous characteristical algorithms are generally similar to those for the synchronous ones.
+In order to formulate these results exactly, let us consider a trial sequence $\{ x^k\} $ generated by an asynchronous characteristic algorithm.
+Assume for the simplification that the number of processors is fixed and equal to \textit{p}.
+Let us associate with the trial sequence $p$ subsequences $\{ \vartheta _{s}^{j} \} ,\, 1\le j\le s$, where $\vartheta _{s}^{j} $ is the moment of completing the $s-$th trial by the $j-$th processor.
+Let us consider the values
+	\begin{equation}
+	\label{eq:3_64}
+	\vartheta _{s}^{j} -\vartheta _{s-1}^{j} ,\; 1\le j\le p,s>1.
+	\end{equation}
+
+	Assume that the number of idle processors $\pi \le card(\Sigma _{k} )$, i.e., none of the idle processors is to wait for the appearance of a free interval to execute a new trial.
+If it is possible to neglect the time of the formation of the set  (\ref{eq:3_63}), the value $\vartheta _{s}^{j} -\vartheta _{s-1}^j $ is the execution time of the \textit{s-}th trial by the \textit{j-}th processor.
+However, in general, $\vartheta _{s}^j -\vartheta _{s-1}^{j} $ are the times from the completing of the ($s - 1$)-th trial till the completion of the $s-$th trial by the \textit{j-}th processor and can include the time of the new trial point selection for the $j-$th processor according to Steps 2 - 6 of the asynchronous algorithm.
+
+	Assume that the times  (\ref{eq:3_64}) are bounded from above, i.e., there is a constant \textit{$\vartheta$}  such that
+
+	\begin{equation}
+	\label{eq:3_65}
+	\vartheta _{s}^{j} -\vartheta _{s-1}^{j} \le \tilde{\vartheta },\; 1\le j\le p,s>1.
+	\end{equation}
+
+	In fact, the relation  (\ref{eq:3_65}) ensures any trial in the asynchronous scheme to be completed in a finite time necessarily.
+
+	In this case (if the additional condition  (\ref{eq:3_65}) holds) Theorems \ref{theor:3_1}-\ref{theor:3_4} take place for the asynchronous methods as well; the proofs of those differ from the ones for the synchronous algorithms in minor details.
+
+	The quantitative efficiency enhancement of the asynchronous algorithms as compared to the synchronous ones is demonstrated in Chapter 4 by means of experiments with the test functions being multiextremal essentially.
+
+\begin{thebibliography}{99.}
+\bibitem{3_Gorodetsky} Gorodetsky, S.Y.: Convergence and asymptotic estimates of behavior for one class of search methods.
+In: System Dynamics.
+Dynamics and Control.
+Gorki State University Publishing (1984).
+In Russian
+\bibitem{3_GoroGri} Gorodetsky, S.Y., Grishagin, V.A.: Nonlinear Programming and Multiextremal Optimization, Models and Methods of Finite-Dimensional Optimization, vol.
+2.
+NNGU Press, Nizhni Novgorod, Russia (2007).
+In Russian
+\bibitem{3_GrishaginSergeyevStrongin} Grishagin, V.A., Sergeyev, Y.D., Strongin, R.G.: Parallel characteristical algorithms for solving problems of global optimization.
+J.
+Global Optim.
+\textbf{10}(2), 185–206 (1997)
+\bibitem{3_GrishaginStrongin_EnginCybernetics} Grishagin, V.A., Strongin, R.G.: Optimization of multiextremal functions subject to monotonically unimodal constraints.
+Engineering Cybernetics \textbf{22},  117--122 (1984)
+\bibitem{3_HansenJaumardLu} Hansen, P., Jaumard, B., Lu, S.H.: Global Optimization of Univariate Lipschitz Functions:2.
+New Algorithms and Computational Comparison.
+Math.
+Program.
+\textbf{55}, 273--293 (1989)
+\bibitem{3_Kushner}	Kushner, H.J.: A new method of locating the maximum point of an arbitrary multipeak curve in the presence of noise.
+Transactions of ASME, Ser.
+D.
+Journal of Basic Engineering \textbf{86}, 97--106 (1964)
+\bibitem{3_LucidiPiccioni} Lucidi, S., Piccioni, M.:  Random tunneling by means of acceptance-rejection sampling for global optimization. J. Optim. Theory Appl. \textbf{62}(2), 255--277 (1989)
+\bibitem{3_Pinter}	Pint{\'e}r, J.D.: Global Optimization in Action.: Kluwer Academic Publishers, Dordrecht (1996)
+\bibitem{3_Piyavskij}	Piyavskij, S.A.: An Algorithm for Finding the Absolute Extremum of a Function. USSR Comput. Math. Math. Phys. \textbf{12}(4), 57--67 (1972)
+\bibitem{3_SergeyevLocTun} Sergeyev, Y.D.: An information global optimization algorithm with local tuning. SIAM J. Optim. \textbf{5}(4), 858–870 (1995)
+\bibitem{3_SergGri1994} Sergeyev, Y.D., Grishagin, V.A.: A Parallel Method for Finding the Global Minimum of Univariate Functions. J. Optim. Theory Appl. \textbf{80}(3), 513--536 (1994)
+\bibitem{3_SergeyevGrishaginOMS} Sergeyev, Y.D., Grishagin, V.A.: Sequential and parallel algorithms for global optimization, Optim. Methods Softw. \textbf{3},  111--124 (1994)
+\bibitem{3_SergMukhKvasLera} Sergeyev, Y.D., Mukhametzhanov, M.S., Kvasov, D.E., Lera, D.: Derivative-free local tuning and local improvement techniques embedded in the univariate global optimization.
+J.
+Optim.
+Theory Appl.
+171(1), 186–208 (2016)
+\bibitem{3_Shubert} Shubert, B.O.: A sequential method seeking the global maximum of a function. SIAM J. Numer. Anal. \textbf{9}(3), 379–388 (1972)
+\bibitem{3_StrMonRus}	Strongin, R.G.: Numerical Methods in Multiextremal Problems (Information-Statistical Algorithms). Nauka, Moscow (1978). In Russian
+\bibitem{3_StrMarkin} Strongin, R.G., Markin, D.L.: Minimization of multiextremal functions with nonconvex constraints. Cybernetics 22, 486–493 (1986)
+\bibitem{3_StrSergMon2000} Strongin, R.G., Sergeyev, Y.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000)
+\bibitem{3_Zilinskas1975} $\check{Z}$ilinskas, A.: One-step Bayesian method for the search of the optimum of one-variable functions. Cybernetics 1, 139--144 (1975)
+\end{thebibliography}
 

### 2017-10-06 2017/ParOptBook/Chapter5.tex
Chapters 3 and 5 have been included. There is a question about cross-references between chapters.

--- 2017-09-01 2017/ParOptBook/Chapter7.tex
@@
 %\usepackage{type1cm}        % activate if the above 3 fonts are
-                             % not available on your system
-%
+                            %% not available on your system
+%%
 %\usepackage{makeidx}         % allows index generation
 %\usepackage{graphicx}        % standard LaTeX graphics tool
-%                             % when including figure files
+                             %% when including figure files
 %\usepackage{multicol}        % used for the two-column index
-%\usepackage[bottom]{footmisc}% places footnotes at page bottom
-%
-%% see the list of further useful packages
-%% in the Reference Guide
+%\usepackage[bottom]{footmisc}
+%\usepackage{amsmath}
+%\usepackage{multirow}
+%\newcommand{\extr}{extr}
 %
 %\makeindex             % used for the subject index
-%                       % please use the style svind.ist with
-%                       % your makeindex program
+                       %% please use the style svind.ist with
+                       %% your makeindex program
 %
@@
 
-\title{Software Systems for Global Optimization }
+\title{Sequential and parallel algorithms for multidimensional multiextremal optimization based on the nested  scheme of dimensionality reduction}
+\titlerunning{Nested  scheme of dimensionality reduction}
 % Use \titlerunning{Short Title} for an abbreviated version of
 % your contribution title if the original one is too long
-\author{Name of First Author and Name of Second Author}
+\author{Vladimir A.Grishagin and Ruslan A.Israfilov}
 % Use \authorrunning{Short Title} for an abbreviated version of
 % your contribution title if the original one is too long
-\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
-\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
-%
-% Use the package "url.sty" to avoid
-% problems with special characters
-% used in your e-mail or web address
+\institute{Vladimir A.Grishagin \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia   \email{vagris@unn.ru}
+\and Ruslan A.Israfilov \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia 
+\email{ruslan@israfilov.com}}
 %
@@
 
-\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\abstract{The dimensionality reduction schemes occupy an important place among the complexity reduction approaches in the multiextremal optimization problems.
+In the present chapter, one of such schemes, namely, so called nested optimization scheme is considered.
+This scheme allows one to replace solving a multidimensional problem by solving a family of one-dimensional subproblems nested recursively.
+Along with a general reduction scheme, its implementation in combination with various global search algorithms is considered.
+The nested scheme has a considerable potential of parallelization since various one-dimensional parallel optimization algorithms, in particular, the characteristical ones can be applied for solving the internal one-dimensional subproblems.
+The asynchronism of the parallel computations originating from the recursive nesting of the subproblems is an integral part of the scheme to be considered.}
 
-\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\section{General nested optimization scheme}
+\label{sec:5_1}
+Let us consider a nonlinear programming problem in the form 
+\begin{equation}
+\label{eq:5_1}
+\varphi(y)\rightarrow min,y\in Q\subseteq R^N,
+\end{equation}
+\begin{equation}
+\label{eq:5_2}
+Q=\{y\in D:g_j(y)\leq 0,\:1\leq j\leq m\},
+\end{equation}
+\begin{equation}
+\label{eq:5_3}
+D=\{y\in R^N:a_i\leq y\leq b_i,\:1\leq i\leq N\}.
+\end{equation}
 
-\section{Section Heading}
-Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.
+Assume that all constraint functions $g_j(y),\:1\leq j\leq m$,  are continuous and the domain $D$  is  bounded that provides the compactness of the feasible domain $Q$.
+
+Let us introduce a continuous function $G(y)$  defined in the domain $D$  such that
+\begin{equation}
+\label{eq:5_4}
+  \begin{cases}
+    G(y)\leq 0, & y\in Q, \\
+    G(y)>0, & y\notin Q.
+  \end{cases}
+\end{equation}
+
+For example, the following functions can be taken as $G(y)$ :
+\begin{equation}
+\label{eq:5_5}
+G(y)=\max\{g_1{y},\;g_2(y),\:\ldots\:,g_m(y)\}
+\end{equation}
+or
+\begin{equation}
+\label{eq:5_6}
+G(y)=\max\{0;\:g_1{y},\;g_2(y),\:\ldots\:,g_m(y)\}.
+\end{equation}
+The latter function equals to zero at all points of the domain $Q$.
+
+The following notations
+\begin{equation}
+\label{eq:5_7}
+u_i=(y_1,\ldots,\:y_i),\;v_i=(y_{i+1},\ldots,y_N),
+\end{equation}
+allow one to write down the vector $y$ as a pair $y=(u_i,\;v_i)$  for $1\leq i\leq N$. Assume that $y=v_0$  if $i=0$  and $y=u_n$  for $i=N$.
+
+Let us introduce  \textit{sections} of the domain $Q$ as
+\begin{equation}
+\label{eq:5_8}
+S_1=Q,\ S_{i+1}(u_i)=\{v_i\in R^{N-i}:(u_i,v_i)\in Q\},\:1\leq i\leq N-1,
+\end{equation}
+and \textit{projections}
+\begin{equation}
+\label{eq:5_9}
+\Pi_{i+1}(u_i)=\{y_{i+1}\in R^1:\exists (u_{i+1},v_{i+1})\in S_{i+1}(u_i)\},\:1\leq i\leq N-1,
+\end{equation}
+of the sections $S_{i+1}(u_i)$ onto the axis $y_{i+1}$.
+
+Let us define recursively a family of functions
+\begin{equation}
+\label{eq:5_10}
+  \begin{cases}
+    G^N(y)\equiv G(y), \\
+    G^i(u_i)=\min\{G^{i+1}(u_i,y_{i+1}:y_{i+1}\in [a_{i+1},b_{i+1}]\},\;1\leq i\leq N-1,
+  \end{cases}
+\end{equation}
+over corresponding projections 
+\begin{equation}
+\label{eq:5_11}
+D_i=\{u_i\in R^i:y_j\in[a_j,b_j],\:1\leq j\leq i\}
+\end{equation}
+of the domain $D$  from (\ref{eq:5_3}) onto the coordinate axes $y_1,\ldots,y_i$. Moreover, by definition, $D_N=D$.
+
+Because of the continuity of the function $G^N(y)\equiv G(y)$ and of the compactness of the domain $D$ the function  $G^{N-1}(u_{N-1})$ exists and is continuous in  $D_{N-1}$.
+This fact causes the existence and the continuity of the function  $G^{N-2}(u_{N-2})$ in  $D_{N-2}$ and further determines the existence and the continuity of all functions from family (\ref{eq:5_10}) in corresponding domains (\ref{eq:5_11}).
+
+The following lemma takes place.
+\begin{lemma} 
+\label{lem:5_1}
+\begin{equation}
+\label{eq:5_12}
+G^i(u_i)=\min\{G(u_i,v_i):y_j\in [a_j,b_j],\:i+1\leq j\leq N\}.
+\end{equation}
+\end{lemma}
+\begin{proof}
+As a consequence of (\ref{eq:5_10}) the assertion (\ref{eq:5_12}) is equivalent to
+\begin{equation}
+\label{eq:5_13}
+\min_{y_{i+1}\in [a_{i+1},b_{i+1}]}\ldots \min_{y_N\in [a_N,b_N]}G(u_i,v_i)=\min\{G(u_i,v_i):y_j\in [a_j,b_j],i+1\leq j\leq N\}.
+\end{equation}
+
+Because of the continuity of $G(y)$, for any $u_i\in D_i$ there exists $\bar{v}_i=(\bar{y}_{i+1},\ldots,\bar{y}_N)$ such that $\bar{y}_j\in [a_j,b_j],\:i+1\leq j\leq N,$ and
+\begin{displaymath}
+\min_{y_{i+1}\in [a_{i+1},b_{i+1}]}\ldots \min_{y_N\in [a_N,b_N]}G(u_i,v_i)=G(u_i,\bar{v}_i)
+\end{displaymath}
+from where it follows that the left-hand side of the equality  (\ref{eq:5_13}) is greater or equal to the right-hand side.
+
+Now the subject of our interest is the reverse inequality.
+Because of the continuity of $G(y)$ and of the compactness of the domain $D$  there exists a vector $v_i^*=(y_{i+1}^*,\ldots ,y_N^*)$  such that
+\begin{displaymath}
+G(u_i,v_i^*)=\min\{G(u_i,v_i):y_j\in [a_j,b_j],\:i+1\leq j\leq N\}.
+\end{displaymath}
+
+According to (\ref{eq:5_10}), we have
+\begin{displaymath}
+G^{N-1}(u_i,y_{i+1}^*,\ldots ,y_{N-1}^*)=\min\{G(u_i,y_{i+1}^*,\ldots,y_{N-1}^*,y_N):y_N\in [a_N,b_N]\}\leq G(u_i,v_i^*),
+\end{displaymath}
+\begin{displaymath}
+G^{N-2}(u_i,y_{i+1}^*,\ldots ,y_{N-2}^*)=\min\{G^{N-1}(u_i,y_{i+1}^*,\ldots,y_{N-2}^*,y_{N-1}):y_{N-1}\in [a_{N-1},b_{N-1}]\}\leq 
+\end{displaymath}
+\begin{displaymath}
+\leq G^{N-1}(u_i,y_{i+1}^*,\ldots,y_{N-2}^*,y_{N-1}^*)\leq G(u_i,v_i^*),
+\end{displaymath}
+$\ldots$
+\begin{displaymath}
+G^i(u_i)=\min\{G^{i+1}(u_i,y_{i+1}):y_{i+1}\in [a_{i+1},b_{i+1}]\}\leq G^{i+1}(u_i,y_{i+1}^*)\leq G(u_i,v_i^*).
+\end{displaymath}
+This chain of inequalities finalizes the proof.
+\end{proof}
+
+Let us introduce the projections 
+\begin{equation}
+\label{eq:5_14}
+Q_i=\{u_i\in R^i:\exists (u_i,v_i)\in Q\},\:1\leq i\leq N,
+\end{equation}
+of the domain $Q$ onto the coordinate axes $y_1,\ldots,y_i$.
+\begin{lemma} 
+\label{lem:5_2}
+Projection (\ref{eq:5_14}) can be presented in the equivalent form 
+\begin{equation}
+\label{eq:5_15}
+Q_i=\{u_i\in R^i:G^i(u_i)\leq 0\}.
+\end{equation}
+\end{lemma}
+\begin{proof}
+Let  (\ref{eq:5_14}) be fulfilled, i.e., for some $u_i$  there exists $v_i^*$  such that $(u_i,v_i^*)\in Q$. Then, $G(u_i,v_i^*)\leq 0$ , i.e., as a consequence of Lemma \ref{lem:5_1},
+\begin{displaymath}
+G^i(u_i)=\min\{G(u_i,v_i):y_j\in[a_j,b_j],\:i+1\leq j\leq N\}\leq G(u_i,v_i^*)\leq 0,
+\end{displaymath}
+and (\ref{eq:5_15}) is true.
+
+Now assume that $G^i(u_i)\leq 0$  for some $u_i\in Q_i$. However, according to Lemma \ref{lem:5_1}, a vector $v_i^*$  exists such that $G^i(u_i)=G(u_i,v_i^*)\leq 0$ , i.e., $(u_i,v_i^*)\in Q$.
+
+Lemma has been proved.
+\end{proof}
+\begin{lemma} 
+\label{lem:5_3}
+The definition of the projection (\ref{eq:5_9}) is equivalent to the form 
+\begin{equation}
+\label{eq:5_16}
+\Pi_{i+1}(u_i)=\{y_{i+1}\in [a_{i+1},b_{i+1}]:G^{i+1}(u_i,y_{i+1})\leq 0\}.
+\end{equation}
+\end{lemma}
+\begin{proof}
+First of all, note that because of $Q\subseteq D$, for all $y\in Q$ the double inequalities $a_j\leq y_j\leq b_j$,  , hold necessarily.
+Let $y_{i+1}$  be such that $G^{i+1}(u_i,y_{i+1})\leq 0$ .
+Then there exists $v_{i+1}^*$  such that $G(u_i,y_{i+1},v_{i+1}^*)=G^{i+1}(u_i,y_{i+1})\leq 0$, i.e., $(y_{i+1}, v_{i+1}^*)\in S_{i+1}(u_i)$.
+Consequently, $y_{i+1}$  belongs to the projection $\Pi_{i+1}(u_i)$  in the sense of the definition (\ref{eq:5_9}).
+
+Next, for any $y_{i+1}$  satisfying (\ref{eq:5_9}) there exists $v_{i+1}^*$  such that $(y_{i+1},v_{i+1}^*)\in S_{i+1}(u_i)$.
+Then, the vector $(u_i,y_{i+1},v_{i+1}^*)\in Q$, i.e., $G^{i+1}(u_i,y_{i+1})\leq G(u_i,y_{i+1},v_{i+1}^*)\leq 0$.
+
+The equivalence of (\ref{eq:5_9}) and (\ref{eq:5_16})  has been proved.
+\end{proof}
+Now consider  a continuous function $\varphi(y)$. Setting $\varphi^N(y)\equiv \varphi(y)$ by definition, let us construct a family of functions 
+\begin{equation}
+\label{eq:5_17}
+\varphi^i(u_i)=\min\{\varphi^{i+1}(u_i,y_{i+1}):y_{i+1}\in \Pi_{i+1}(u_i)\},\:1\leq i\leq N-1,
+\end{equation}
+defined over the corresponding projections $Q_i$  from (\ref{eq:5_14}). Then, according to \cite{5_CarrHowe} and \cite{5_StrSergMon2000} the basic relation of the nested optimization scheme 
+\begin{equation}
+\label{eq:5_18}
+\min_{y\in Q}\varphi(y)=\min_{y_1\in \Pi_1}\min_{y_2\in \Pi_2(u_1)}\ldots\min_{y_N\in \Pi_N(u_{N-1})}\varphi(y).
+\end{equation}
+takes place.
+
+As it follows from (\ref{eq:5_18}), in order to solve the problem (\ref{eq:5_1})--(\ref{eq:5_3}) it is sufficient to solve the one-dimensional problem 
+\begin{equation}
+\label{eq:5_19}
+\varphi^1(y_1)\rightarrow \min,\:y_1\in \Pi_1\subseteq R^1,
+\end{equation}
+\begin{displaymath}
+\Pi_1=\{y_1\in [a_1,b_1]:G^1(y_1)\leq 0\}.
+\end{displaymath}
+
+According to (\ref{eq:5_17}), each calculation of the function $\varphi^1(y_1)$  at some fixed point $y_1\in\Pi_1$   requires solving  the one-dimensional problem 
+\begin{displaymath}
+\varphi^2(y_1,y_2)\rightarrow \min,\:y_2\in \Pi_2(y_1)\subseteq R^1,
+\end{displaymath}
+\begin{displaymath}
+\Pi_2(y_1)=\{y_2\in [a_2,b_2]:G^2(y_1,y_2)\leq 0\}.
+\end{displaymath}
+This problem is a one-dimensional minimization problem with respect to $y_2$  since $y_1$  is fixed.
+
+In turn, each calculation of the function $\varphi^2(y_1,y_2)$  with fixed $y_1,y_2$ generates solving the one-dimensional problem
+\begin{displaymath}
+\varphi^3(u_2,y_3)\rightarrow \min,\:y_3\in \Pi_3(u_2)\subseteq R^1,
+\end{displaymath}
+etc., up to solving the univariate problem
+
+\begin{equation}
+\label{eq:5_20}
+\varphi^N(u_{N-1},y_N)\rightarrow \min,\:y_N\in \Pi_N(u_{N-1})\subseteq R^1,
+\end{equation}
+where $u_{N-1}$  is fixed.
+
+Thus, solving the problem (\ref{eq:5_1})--(\ref{eq:5_3}) can be reduced to solving a family of «nested» one-dimensional subproblems 
+\begin{equation}
+\label{eq:5_21}
+\varphi^i(u_{i-1},y_i)\rightarrow \min,\:y_i\in \Pi_i(u_{i-1}),
+\end{equation}
+where the fixed vector $u_{i-1}\in Q_{i-1}$.
+
+Solving initial multidimensional problem (\ref{eq:5_1})--(\ref{eq:5_3})  through solving the system of the one-dimensional subproblems (\ref{eq:5_21}) is called \textit{the nested optimization scheme}.
+
+On the base of the nested scheme (\ref{eq:5_18}) many multidimensional optimization algorithms have been designed (see, for example, \cite{5_CarrHowe, 5_Evtushenko, 5_GerGriIsr, 5_GriIsrAIP, 5_GriIsrCEUR, 5_GrishaginStrongin_EnginCybernetics, 5_Piyavskij, 5_SergGriJCAA, 5_ShiOlaf, 5_StrSergMon2000, 5_vanDam}).
+This scheme has essential potential for constructing new optimization techniques by means of combining with different one-dimensional algorithms for solving the subproblems (\ref{eq:5_21}) (for example, characteristical methods \cite{5_GrishaginSergeyevStrongin} including information    algorithms \cite{5_SergeyevLocTun, 5_Sergeyev1998, 5_SergeyevGrishaginOMS, 5_SergMukhKvasLera, 5_StrMonRus, 5_Str1973, 5_StrMarkin, 5_StrSergMon2000}, methods by Piyavskij and Shubert \cite{5_Piyavskij, 5_Shubert} and Kushner \cite{5_Kushner}, Bayesian algorithms \cite{5_Locatelli, 5_Mockus1980, 5_Mockus1988, 5_MockusEddy_et_al, 5_Zilinskas1975, 5_Zilinskas1981, 5_Zilinskas1985}, etc.)
+
+As an illustration, let us consider an example of the maximization of a two-dimensional multiextremal function from the class (\ref{eq:3_518}) using the nested optimization scheme.
+The surface and level curves of the objective function are shown in Fig.
+5.1 and in Fig.
+\ref{fig:5_2}, respectively.
+The coordinates of the search trials of the nested scheme are shown in Fig.
+5.2 also.
+For this function the global maximum value 13.51 is located at the point (0.603, 0.408).
+The method AGS (\ref{eq:2_38})--(\ref{eq:2_40}) was employed for solving the one-dimension subproblems (\ref{eq:5_21}).
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.1\linewidth]{figures/figure_5_2.pdf}
+\caption{Level curves and trial points}
+\label{fig:5_2}    
+\end{figure}
+
+The nested scheme (\ref{eq:5_18})--(\ref{eq:5_21}) may be generalized in the following way.
+Let us consider for simplicity a problem (\ref{eq:5_1}) without functional constraints, i.e., $Q=D$, and present the feasible domain  as the direct product of subdomains with smaller dimensions
+\begin{displaymath}
+D=D_1\times D_2\times\:\ldots\: D_s,\:1\leq s<N,
+\end{displaymath}
+where the subdomains $D_i,\:1\leq i\leq s,$ are hyperparallelepipeds (\ref{eq:5_3}) of corresponding dimensions $d_i$ such that $1\leq d_i<N$ and $\sum_{i=1}^s{d_i}=N$.
+
+In this case (\ref{eq:5_18}) can be transformed to the form
+\begin{displaymath}
+\min_{y\in Q}\varphi(y)=\min_{z_1\in D_1}\min_{z_2\in D_2}\ldots\min_{z_s\in D_s}\varphi(y),
+\end{displaymath}
+where $z_1=u_{\sigma(1)},\:z_i=(y_{\sigma{(i-1)}}+1,\ldots, y_{\sigma(i)}),\:2\leq i\leq s,$ and $\sigma(j)=\sum_{i=1}^j{d_i},\:1\leq j\leq s$.
+
+Then, instead of the family (\ref{eq:5_21}) we can solve the set of nested subproblems 
+\begin{displaymath}
+\varphi(u_{\sigma(i-1)},z_i)\rightarrow \min,\;z_i\in D_i,
+\end{displaymath}
+which can be multidimensional with lesser dimensions.
+Such approach was developed (see \cite{5_BarkGer2014, 5_BarkGerLeb, 5_BarkLeb, 5_SysoyevBarkGerLeb}) in combination of the optimization algorithms based on the Peano space-filling curves for solving the nested multidimensional subproblems.
+
+Moreover, in the family (\ref{eq:5_21}) for several functions $\varphi^i(u_i)$ one can use the operation of taking maximum instead of minimum.
+In this case it is possible to calculate complex minmax (or maxmin) expressions.
+
+\section{Properties of one-dimensional subproblems in the nested optimization scheme}
+\label{sec:5_2}
+While analyzing univariate subproblems (\ref{eq:5_21}) of the nested scheme, two main problems arise:
+\begin{description} [a)]
+\item [a)] {the necessity to construct the feasible domains $\Pi_i(u_{i-1})$  for the one-dimensional search;}
+\item [b)] {it is required to provide the minimization of the one-dimensional functions $\varphi^i(u_{i-1},y_i)$  in the domains  $\Pi_i(u_{i-1})$.}
+\end{description}
+
+The structure and the complexity of the projections $\Pi_i(u_{i-1})$  are completely determined by the complexity of the multidimensional feasible domain $Q$.
+The complexity of the second problem depends on the characteristics of the functions $\varphi^i(u_i)$ .
+These characteristics are influenced by the properties of the objective function $\varphi(y)$ as well as by the features of the search domain $Q$ defined, in turn, by constraints (\ref{eq:5_2}) and (\ref{eq:5_3}).
+
+\subsection {Structure of the feasible domains of the one-dimensional search}
+\label{subsec:5_2_1}
+The results of Lemma \ref{lem:5_3}, which has established the equivalence of Definition (\ref{eq:5_9}) and representation (\ref{eq:5_16}) enable to analyze the structure of the domains $\Pi_i(u_{i-1})$.
+In fact, (\ref{eq:5_16}) provides a constructive apparatus for building the domains $\Pi_i(u_{i-1})$  by means of finding the domains of the non-positivity for the functions $G^i(u_{i-1},y_i)$.
+
+Since the function $G(y)$ is assumed to be continuous in the domain $D$, the functions $G^i(u_i)$ are continuous with respect to $u_i\in D_i$  from (\ref{eq:5_11}) and, therefore, with regard to $y_i\in [a_i,b_i]$  as well.
+Then, at fixed $u_{i-1}\in D_{i-1}$  each one-dimensional problem (\ref{eq:5_21}) can be rewritten in a unified form
+\begin{equation}
+\label{eq:5_22}
+  \begin{cases}
+    \bar{\varphi}(x)\rightarrow\min,\;x\in\bar{Q}\subseteq R^1, \\
+    \bar{Q}=\{x\in[a,b]:g(x)\leq 0\},
+  \end{cases}
+\end{equation}
+where the function $g(x)$ is continuous.
+
+The continuity of the constraint $g(x)$ allows one to state that the feasible domain $\bar{Q}$  can be presented as a system of intervals
+\begin{equation}
+\label{eq:5_23}
+\bar{Q}=\bigcup_{j=1}^q{[a^j,b^j]},
+\end{equation}
+in each of which the function $g(x)$ is non-positive.
+
+As an example, let us consider the possible cases of the behavior of a continuous function generating the domains of the non-positivity presented in Fig.
+\ref{fig:5_3}.
+The corresponding intervals  are marked with  the signs $\times$  on the abscissa axis including the points of tangency.
+The most right-side interval marked with $\times$ corresponds to the case when the function equals to zero over the interval of non-zero length.
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.1\linewidth]{figures/figure_5_3.pdf}
+\caption{Domains of the non-positivity of a continuous function}
+\label{fig:5_3}    
+\end{figure}
+In the system (\ref{eq:5_23}) the number of intervals $q$ can be infinite. As an example of such situation one can consider within the closed interval $[0,1]$ the function
+\begin{displaymath}
+g(x)=
+  \begin{cases}
+    x\sin(\frac{1}{x}), & x>0,\\
+    0, & x=0.
+  \end{cases}
+\end{displaymath}
+
+Thus, in the case of a continuous function $G(y)$, the projection $\Pi_i(u_{i-1}$ determined as  (\ref{eq:5_16}) is a set of the kind (\ref{eq:5_23}), i.e.,
+\begin{equation}
+\label{eq:5_24}
+\Pi_i(u_{i-1})=\bigcup_{j=1}^{q_i}{[a_i^j,b_i^j]},
+\end{equation}
+where, in general, the number of the intervals $q_i$  and their bounds $a_i^j$, $b_i^j$, $1\leq j\leq q_i,$ depend on the vector $u_{i-1}$ , i.e.,
+\begin{equation}
+\label{eq:5_25}
+q_i=q_i(u_{i-1}),\:a_i^j=a_i^j(u_{i-1}),\:b_i^j=b_i^j(u_{i-1}).
+\end{equation}
+
+If the domain $Q$  is such that it is possible to obtain the explicit (analytical) expressions for the values $q_i,a_i^j,b_i^j$   as  functions of $u_{i-1})\in Q_{i-1}$  for all $1\leq i\leq N$, then the domain $Q$ is called \textit{the domain with the computable bounds}.
+In particular, it is possible if all the roots of the functions $G^i(u_{i-1},y_i)$  with regard to the variables $y_i$  can be analytically found.
+\begin{example}
+\label {exam:5_1}
+As the problem to be investigated  the problem (\ref{eq:5_1})--(\ref{eq:5_3}) with 
+\begin{equation}
+\label{eq:5_26}
+  \begin{cases}
+    \varphi(y)=y_1^2+y_2^2,\\
+   Q=\{y\in D:(y_1-1)^2+(y_2-1)^2-1\leq 0\}, \\
+	D=\{y\in R^2:0\leq y_1,y_2\leq 2\}
+  \end{cases}
+\end{equation}
+is considered.
+The feasible domain is marked with gray in Fig.
+\ref{fig:5_4}.
+Also, the level curves of the objective function are shown there.
+In this problem, the function $G^2(y)=(y_1-1)^2+(y_2-1)^2-1$.
+It is evident that the function  $G^1(y_1)=\min\{G^2(y_1,y_2):y_2\in [0,\;2]\}=(y_1-1)^2-1$.
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.1\linewidth]{figures/figure_5_4.pdf}
+\caption{Feasible domain and level curves of the objective function}
+\label{fig:5_4}    
+\end{figure}
+
+According to (\ref{eq:5_16}), the domains where the functions $G^1(y_1)$  for $y_1$  and $G^2(y_1,y_2)$  for $y_2$  are not positive determine the projections $\Pi_1$  and $\Pi_2(y_1)$ , respectively.
+The domain bounds are the roots of these functions in the interval $[0,2]$.
+For the function $G^1(y_1)$  these roots are $0$ and $2$.
+Therefore, $\Pi_1=[0,\:2]$.
+
+The function $G^2(y)=(y_2-1)^2-\alpha^2$, where $\alpha=\sqrt{1-(y_1-1)^2}\leq 1$,  has the roots $1\pm \alpha$ within the interval $[0,2]$ and  it is not positive between these roots. As a result, 
+\begin{equation}
+\label{eq:5_27}
+\Pi_2(y_1)=[1-\sqrt{1-(y_1-1)^2},1+\sqrt{1-(y_1-1)^2}].
+\end{equation}
+
+Thus, we can write down the bounds of projections (\ref{eq:5_24}) in an explicit form. Thereby, the domain $Q$ from (\ref{eq:5_26}) is a domain with computable bounds (\ref{eq:5_25}).
+
+The function $\varphi(y)=y_1^2+y_2^2$  is an increasing one for $y_2$  in the interval $[0,2]$. Consequently, it reaches its maximum at the point $1-\alpha$  in the domain (\ref{eq:5_27}), therefore,
+\begin{displaymath}
+\varphi^1(y_1)=y_1^2+(1-\sqrt{1-(y_1-1)^2})^2=1+2y_1-2\sqrt{1-(y_1-1)^2}.
+\end{displaymath}
+
+The first derivative $(\varphi^1(y_1))'=2+\frac{2(y_1-1)}{\sqrt{1-(y_1-1)^2}}$ has a unique root  $y_1^*=1-\frac{1}{\sqrt{2}}$ in the interval  $[0,2]$. 
+In addition, the second derivative $(\varphi^1(y_1))''=\frac{2}{(1-(y_1-1)^2)^{3/2}}>0$, therefore, the point $y_1^*=1-\frac{1}{\sqrt{2}}$  gives the minimum value of the function $\varphi^1(y_1)$  equal to $3-2\sqrt{2}$.
+This value is the sought minimum of the function $\varphi(y)$ in the domain $Q$.
+
+In order to obtain the coordinate $y_2^*$, which determines the minimum point along with $y_1^*$, let us consider the function $\varphi^2(y_1^*,y_2)$ and find its minimum in the domain $\Pi_2(y_1^*)=[1-\frac{1}{\sqrt{2}},1+\frac{1}{\sqrt{2}}]$, which is obviously achieved at the point $y_2^*=1-\frac{1}{\sqrt{2}}$.
+
+So, the vector $y^*=(1-\frac{1}{\sqrt{2}},1-\frac{1}{\sqrt{2}})$  providing the minimum value $\varphi^*=3-2\sqrt{2}$ of the objective function is the solution of the multidimensional problem (\ref{eq:5_26}).
+The optimum point is marked with a dark cross in Fig.
+\ref{fig:5_4}.
+\end{example}
+
+In the example considered above, we have constructed the one-dimensional search domain bounds  analytically by finding the domains of the non-positivity of the corresponding functions $G^i(u_i)$ .
+At the same time, one can specify a more visual «geometrical» method for the construction of the projections $\Pi_{i+1}(u_i)$.
+This method follows from definitions (\ref{eq:5_8}) and (\ref{eq:5_9}) and consists in the construction of the sections of the domain $Q$  by planes $u_i=const$.
+After that, it is necessary to find the bounds of these sections for the coordinate $y_{i+1}$.
+
+In this regard, the following is noteworthy.
+The computation of the global minimum according to (\ref{eq:5_18}) is analogous to the procedure of computation of a multidimensional integral of the function $\varphi(y)$ over the domain $Q$ by means of the reduction to computing  the recurring (nested) one-dimensional integrals.
+Here, the domains of the one-dimensional integration are just the corresponding projections $\Pi_i(u_{i-1})$.
+
+As an illustration the example \ref{exam:5_2} can be taken. 
+\begin{example}
+\label{exam:5_2}
+Assume that the domain of the optimization (or of the integration) is defined as
+\begin{displaymath}
+Q=\{y\in R^2:-4\leq y_1,y_2\leq 4, y_1^2+y_2^2\leq 4,y_1^2+y_2^2\geq 1\}.
+\end{displaymath}
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\linewidth]{figures/figure_5_5.pdf}
+\caption{Sections and projections}
+\label{fig:5_5}    
+\end{figure}
+In Fig.
+\ref{fig:5_5} this domain is marked with gray.
+The left-side line $y_1=const$  generates the section $S_1(y_1)$  as the intersection of the line with the domain $Q$, and the projection  of $S_1(y_1)$  onto the axis $y_2$  determines the projection $\Pi_2(y_1)$ .
+This projection and the projection $\Pi_1$  are marked with the bold lines in Fig.
+\ref{fig:5_5}.
+
+These geometrical considerations allow constructing the necessary projections of the problem as 
+\begin{displaymath}
+\Pi_1=[-2,\:2],
+\end{displaymath}
+\begin{displaymath}
+\Pi_2(y_1)=
+  \begin{cases}
+    [-\sqrt{4-y_1^2}, \sqrt{4-y_1^2}],\ y_1\in [-2,\:-1]\bigcup [1,\:2],  \\
+    [-\sqrt{4-y_1^2}, -\sqrt{1-y_1^2}]\bigcup [\sqrt{1-y_1^2}, \sqrt{4-y_1^2}],y_1\in [-1,\:1].                      	
+  \end{cases}
+\end{displaymath}
+\end{example}
+
+However, finding all roots of a continuous function is a difficult problem, unsolvable analytically, as a rule.
+In this case, one can try to find these roots numerically.
+Let us consider, for example, typical problem (\ref{eq:5_22}).
+In order to find all roots of a continuous function $g(x)$ one can propose  numerical solving  the equivalent optimization problem
+\begin{displaymath}
+\left|g(x)\right|\rightarrow \min,x\in [a,b],
+\end{displaymath}
+which roots of $g(x)$ are the global minima points in. To solve this problem, the characteristical global search algorithms providing the convergence to all global minimum points might be applied.
+
+Another approach to accounting the constraints in the optimization problems (the index method), which does not require to solve any auxiliary problems  is considered in Chapter 4.
+
+The case $Q=D$  when the constraint functions (\ref{eq:5_2}) are absent is an important particular case of the problem (\ref{eq:5_1})--(\ref{eq:5_3}).
+In this case, $G(y)\equiv 0$ in the domain $D$, and it follows from (\ref{eq:5_12}) that $G^i(u_i)\equiv 0$,$1\leq i\leq N$.
+Then, according to (\ref{eq:5_16}),
+\begin{displaymath}
+\Pi_i(u_{i-1})=[a_i,b_i],
+\end{displaymath}
+where $a_i,b_i$ are constants.
+
+The optimization domains (\ref{eq:5_2}) can generate projections $\Pi_i(u_{i-1})$  as a single interval for some particular properties of the constraint functions as well.
+These properties are the convexity of the functions $g_j(y),\:1\leq j\leq m$, providing the convexity of the feasible domain $Q$ and the monotonous unimodality of the constraint functions \cite{5_GrishaginStrongin_EnginCybernetics}, which is more general than the concept of convexity.
+In general case, monotonous unimodal constraints can generate non-convex domains $Q$.
+
+In the situation of the monotonous unimodality (or of convexity as a particular case) of the constraints $g_j(y),\:1\leq j\leq m$, the one-dimensional search domains are the intervals 
+\begin{displaymath}
+\Pi_i(u_{i-1})=[a_i^1(u_{i-1}),b_i^1(u_{i-1})],
+\end{displaymath}
+the bounds of which, however, are not constant but depend on the vector $u_{i-1}$.
+\subsection {Properties of the objective functions in one-dimensional subproblems}
+\label{subsec:5_2_2}
+
+The objective function in a subproblem (\ref{eq:5_21}) is the function $\varphi^i(u_{i-1},y_i)$  with fixed $u_{i-1}$ .
+The character of the dependence of the function $\varphi^i$  on the variable $y_i$  is crucial while solving the subproblems (\ref{eq:5_21}).
+
+Let us consider a class of problems, in which the function $\varphi(y)$ is separable, i.e.,
+\begin{equation}
+\label{eq:5_28}
+\varphi(y)=\sum_{i=1}^N{\varphi_i(y_i)},
+\end{equation}
+and the constraints $g_j(y)$  are absent, i.e., $Q=D$.
+
+Then, as it follows from (\ref{eq:5_18}) and (\ref{eq:5_28}), 
+\begin{displaymath}
+\min_{y\in Q}\varphi(y)=\sum_{i=1}^N{\min_{y_i\in [a_i,b_i]}\varphi_i(y_i)},
+\end{displaymath}
+i.e., to solve the multidimensional problem it is necessary to solve $N$ independent one-dimensional subproblems.
+For this problem class, the complexity increases linearly with increasing dimensionality.
+
+Now assume that the function $\varphi(y)$ satisfies the Lipschitz condition with a constant $L>0$ in the domain $Q$, i.e., for any $y',y''\in Q$
+\begin{equation}
+\label{eq:5_29}
+\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,
+\end{equation}
+where $\left\|\bullet\right\|$ denotes the Euclidean norm.
+The obvious question in this case is whether the functions $\varphi^i(u_i)$  satisfy the Lipschitz condition.
+Generally speaking, it takes place not always.
+
+Let us consider the following example.
+Assume that two-dimensional objective function $\varphi(y)$  in problem (\ref{eq:5_1})--(\ref{eq:5_3}) satisfies the Lipschitz condition (\ref{eq:5_29}) and the feasible domain
+\begin{displaymath}
+Q=\{y\in R^2:y_1^2+y_2^2-1\leq 0\}.
+\end{displaymath}
+
+The function $\varphi^2(y)\equiv \varphi(y)$ , obviously, is Lipschitzian with the constant $L$ with regard to the coordinate $y_2$.
+However, the function $\varphi^1(y_1)$  does not satisfy the condition (\ref{eq:5_29}).
+As it has been shown in \cite{5_StrMonRus}, this function satisfies the generalized Lipschitz condition (H{\"o}lder condition) in the metric $\rho(y_1',y_1'')=\sqrt{\left|y_1'-y_1''\right|}$  with the constant $L_1=L(1+\sqrt{2})$ , i.e., the inequality
+\begin{displaymath}
+\left|\varphi(y_1')-\varphi(y_1'')\right|\leq L_1\sqrt{\left|y_1'-y_1''\right|}.
+\end{displaymath}
+
+Sufficient conditions for the Lipschitz property (\ref{eq:5_29})  of functions $\varphi^i(u_i)$  is established by the following theorem (see \cite{5_StrMonRus}).
+\begin{theorem}
+\label{theor:5_1}
+Assume that the function $\varphi(y)$ satisfies the Lipschitz condition (\ref{eq:5_29})  with the constant $L>0$ in a domain $Q$ from (\ref{eq:5_2}) and the bound pairs (\ref{eq:5_27}) are the piecewise linear functions of the kind
+\begin{equation}
+\label{eq:5_30}
+a_{i+1}(u_i)=\max_{1\leq\nu \leq p_{i+1}}\{\alpha_i^\nu u_i+A_i^\nu\},
+\end{equation}
+\begin{equation}
+\label{eq:5_31}
+b_{i+1}(u_i)=\max_{1\leq\nu \leq r_{i+1}}\{\beta_i^\nu u_i+B_i^\nu\},
+\end{equation}
+where  $\alpha_i^\nu u_i$ and $\beta_i^\nu u_i$ are the scalar products of the vectors from $R^i$  and $A_i^\nu, B_i^\nu$ are constants.
+Then the functions $\varphi^i(u_i),u_i\in Q_i,$ satisfy the Lipschitz condition with the constants $L_i,1\leq i\leq N,$
+where
+\begin{displaymath}
+L_N=L,\:L_i=L\sum_{j=i}^{N-1}{(1+\lambda_j)},\:1\leq i\leq N-1,
+\end{displaymath}
+\begin{displaymath}
+\lambda_j=\max\left\{\max_{1\leq \nu\leq p_{i+1}}\left\|a_j^\nu\right\|,\max_{1\leq \nu\leq r_{i+1}}\left\|b_j^\nu\right\| \right\}.
+\end{displaymath}
+\end{theorem}
+
+The proof of the theorem is given in \cite{5_StrMonRus}.
+
+Note that the representation of the boundy pairs in the form (\ref{eq:5_30}) and (\ref{eq:5_31}) takes place if the feasible domain is a convex polyhedron.
+
+If the domain $Q$ is a hyperparallelepiped, all vectors $a_i^\nu,b_i^\nu$ are equal to zero. Therefore, $L_i=L,\:1\leq i\leq N$.
+
+The main conclusion, which follows from the discussion of Lipschitz properties is that the properties of the objective functions of the one-dimensional problems depend essentially not only on the properties of the initial objective function $\varphi(y)$  but also on the shape of the feasible domain $Q$.
+
+In the simplest case, when the domain $Q$ is a hyperparallelepiped all functions $\varphi_i(u_i)$  satisfy Lipschitz condition with the same constant as the function $\varphi(y)$.
+
+If the constraints $g_j(y)$   are piecewise linear convex functions, the Lipschitzness of the functions $\varphi_i(u_i)$  remains but Lipschitz constants for these functions increase in general that worsens the optimization properties of the one-dimensional subproblems.
+
+Finally, the case of the nonlinear constraints can result in the loose of Lipschitzness at all.
+\section{Parallel computations for the nested optimization scheme}
+\label{sec:5_3}
+Applying the parallel methods of the one-dimensional optimization for solving one-dimensional subproblems (\ref{eq:5_21}) in the structure of the nested scheme (\ref{eq:5_18}) enables to obtain a parallel version of the nested scheme with a high degree of variability.
+For example, one can vary the number of processors employed at different levels of the one-dimensional optimization (i.e., during solving the one-dimensional subproblems for different variables ), apply various parallel one-dimensional optimization methods at different levels, etc.
+
+In order to describe the parallel nested scheme, let us introduce a vector of the parallelization degrees
+\begin{equation}
+\label{eq:5_32}
+\pi=(\pi_1,\pi_2,\:\ldots\:,\pi_N)
+\end{equation}
+where $\pi_i,1\leq i\leq N$, is the number of one-dimensional subproblems at the $(i+1)$-th recursion level arising as a result of the parallel iterations at the $i$-th level of the one-dimensional optimization and being solved in parallel.
+The value $\pi_N$  for the coordinate $y_N$  is the number of the parallel trials during the minimization of the function $\varphi^N(y_1,\ldots y_N)\equiv  \varphi(y_1,\ldots y_N)$ with fixed values of $y_1,\ldots, y_{N-1}$, i.e., the number of the objective function values computed in parallel.
+
+In general, the values $\pi_i,1\leq i\leq N$, may depend on various parameters and may be changed in the course of the optimization.
+Here, for simplicity, we will consider the case when all components of the vector (\ref{eq:5_32}) are constant.
+
+The application of the one-dimensional parallel algorithms combined with scheme (\ref{eq:5_18})--(\ref{eq:5_21}) and vector (\ref{eq:5_32}) allows using up to
+\begin{displaymath}
+P=\prod_{i=1}^N{\pi_i}
+\end{displaymath}
+processors working in parallel for solving the problem (\ref{eq:5_1})--(\ref{eq:5_3}).
+Note, the nested scheme generates up to $P/\pi_N$  one-dimensional optimization subproblems being solved simultaneously.
+
+Let us consider, for example, the dimensionality $N=2$. The use of a method with $\pi_1$  parallel trials at one iteration to solve the external univariate problem
+\begin{displaymath}
+\varphi^1(y_1)\rightarrow\min,y_1\in \Pi_1\subseteq R^1,
+\end{displaymath}
+generates  $\pi_1$ internal one-dimensional subproblems
+\begin{displaymath}
+\varphi^2(y_1,y_2)\rightarrow\min,y_2\in \Pi_2(y_1)\subseteq R^1.
+\end{displaymath}
+Each of these subproblems can be solved by a parallel method with  $\pi_2$  trials at iteration.
+It means that at the most internal level corresponding  to $i=N$  in (\ref{eq:5_21}) $P=\pi_1\pi_2$  values of the function $\varphi(y)$  are computed, i.e., one can apply $P$  processors running in parallel for computing these values.
+
+One of possible distributions for   $P=6$ parallel processors is presented in Fig. 
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\linewidth]{figures/figure_5_6.pdf}
+\caption{Distribution of processors}
+\label{fig:5_6}    
+\end{figure}
+
+The nested scheme has a significant potential of the parallelization efficiency by means of the use of the asynchronism. This fact is explained by the following reasons.
+
+Let us assume that we combine the nested scheme with a one-dimensional parallel algorithm, which the termination condition differs from realizing a fixed number of iterations in.
+In this case, the execution time of the trials (the computations of the minimized function) in a parallel iteration at the $i$-th level of the recursion ($1\leq i\leq N-1$)  are different necessarily even if the computation time of the objective function $\varphi(y)$  is independent on the arguments.
+Remind that the evaluation of the objective function at all levels except the last one consists in solving of a new optimization subproblem.
+
+Assume that the synchronous scheme of the parallel iterations of the one-dimensional optimization algorithm is applied at the $i$-th level of recursion, when the iteration is considered to be complete after execution of all $\pi_i$   trials only.
+In this case, at least
+\begin{displaymath}
+\prod_{j=i+1}^N{\pi_j}
+\end{displaymath}
+processors, which have already completed solving their subproblems stay idle until the rest subproblems of the $i$-th level are completed.
+This means that for increasing the productivity of the solving process in the nested scheme it is necessary to use the asynchronous parallel global optimization algorithms, in which the computational resources are used immediately after they become free.
+
+\paragraph{Numerical experiments}
+
+In order to illustrate the theoretical statements, the results of the numerical solving of different multidimensional multiextremal problems are presented below.
+Three methods which realize multidimensional optimization using the nested scheme (\ref{eq:5_18})--(\ref{eq:5_21}) with combination of information univariate algorithms are considered:
+\begin{description} [1)]
+\item [1)] {AGSM– algorithm of global search which solves the one-dimensional subproblems inside of the nested scheme by the sequential core information algorithm AGS;}
+\item [2)] {PSIM – parallel synchronous information algorithm. It uses AGSIP for solving the one-dimensional subproblems (\ref{eq:5_21});}
+\item [3)] {AIPM – asynchronous multidimensional information parallel algorithm applying the asynchronous one-dimensional algorithm proposed in \cite{5_SergGriJCAA} for solving the one-dimensional subproblems.}
+\end{description}
+
+The following assumptions are introduced for the first two series of experiments.
+\begin{description} [i)]
+\item [i)] {The times required for computation of the objective function values are the same for all feasible points.
+This time can be used as a “unit” time or a “quantum” of time  in the search process.
+Note, that this assumption is the worst case for the asynchronous method since it is more suitable for the synchronous parallelization.}
+\item [ii)] {The time for computation of the objective function value is much greater than the one spent for the implementation of the computational operations of the algorithmic scheme.}
+\end{description}
+
+This approach allows abstracting from the architecture of the parallel computational system, on which the experiments were carried out and estimate the «pure» efficiency of the algorithm decision rules.
+
+Let us denote the time of solving the multidimensional problem (measured in the computational quanta) using the nested reduction scheme with application of the asynchronous algorithm AIPM and the parallelization vector (\ref{eq:5_32}) as $T_A(\pi)=T_A(\pi_1,\pi_2,\ldots,\pi_N)$ , the one with the use of the synchronous algorithm PSIM  (in the analogous conditions) as $T_S(\pi)=T_S(\pi_1,\pi_2,\ldots,\pi_N)$ , and the one with the use of the sequential method AGSM as $T_1=T_S(1,1,\ldots,1)=T_A(1,1,\ldots,1)$ .
+The last time is simply equal to the number of the computed values of the function $\varphi(y)$.
+
+Furthermore, the numbers of trials performed by the asynchronous method and by the synchronous one will be denoted as $K_A$  and $K_S$, respectively.
+Using the introduced values, one can define the corresponding time speed-ups achieved owing to the parallelism.
+
+$u_S(\pi)=T_1/T_S(\pi)$ is the speed up of the synchronous algorithm.
+
+$u_A(\pi)=T_1/T_A(\pi)$ is the speed up of the asynchronous algorithm.
+
+As a measure for the comparison of the relative efficiency of the synchronism and the asynchronism one can introduce the coefficient $u(\pi)=u_A(\pi)/u_S(\pi)$.
+
+Note that because of Assumption (i) in the case $\pi=(1,1,\ldots,1,\pi_N)$  we have $u_A(\pi)=u_S(\pi)$ and $u(\pi)=1$.
+
+In the first experiment, the maximization of 100 two-dimensional functions from the class (\ref{eq:3-58}) has been carried out.
+All experiments were carried out with the precision in the termination criterion $\epsilon=0.01$  and with the reliability parameter $r=2$  for solving all one-dimensional subproblems (\ref{eq:5_21}).
+All one-dimensional searches have been started from the same starting points {0.2, 0.3, 0.7, 0.8}.
+
+For the method AGSM, which is a pure sequential one, the average number of trials determined as a result of the maximization of 100 functions of the class (\ref{eq:3-58}) was found to be  $T_1= 415.08$.
+This value can be considered as a measure for the average search time.
+
+The results for the parallel methods are presented in Table \ref{tab:5_1}.
+
+\begin{table}
+\caption{Comparison of the parallel algorithms on the two-dimensional test class}
+\label {tab:5_1}
+\begin{center}
+\begin{tabular}{|c|r|r|r|r|r|}
+\hline
+$q$ & (2,1) & (2,2) & (2,3) & (2,4) & (2,5)   \\
+\hline
+$\ \ $$K_A\ \ $  &  $\ \ \ $409.46 & $\ \ \ $397.96 & $\ \ \ $419.41 & $\ \ \ $456.00 & $\ \ \ $510.60  \\
+\hline
+$K_S$  & 413.89 & 409.24 & 424.08 & 458.60 & 514.20  \\
+\hline
+$T_A$  & 204.73 &  99.49 &  69.90 &  57.00 &  51.06  \\
+\hline
+$T_S$  & 234.52 & 114.63 &  77.79 &  62.46 &  56.89  \\
+\hline
+$u_A$  &   2.03 &   4.17 &   5.94 &   7.28 &   8.13  \\  
+\hline
+$u_S$  &   1.77 &   3.62 &   5.34 &   6.64 &   7.30  \\
+\hline
+$u$    &   1,15 &   1,15 &   1,11 &   1,10 &   1,23  \\
+\hline
+\hline
+$q$ & (3,1) & (3,2) & (3,3) & (3,4) & (3,5)   \\
+\hline
+$K_A$  & 427.23 & 429.18 & 432.90 & 472.20 & 529.95  \\
+\hline
+$K_S$  & 432.40 & 426.16 & 433,83 & 469.44 & 527.90  \\
+\hline
+$T_A$  & 123.41 &  71.53 &  48.10 &  39.35 &  35.33  \\
+\hline
+$T_S$  & 174.38 &  84.90 &  55.95 &  44.63 &  40.62  \\
+\hline
+$u_A$  &   2.91 &   5.80 &   8.63 &  10.55 &  11.75  \\
+\hline
+$u_S$  &   2.38 &   4.89 &   7.42 &   9.30 &  10.22  \\
+\hline
+$u$    &   1.22 &   1.19 &   1.16 &   1.13 &   1.15  \\
+\hline
+\hline
+$q$ & (4,1) & (4,2) & (4,3) & (4,4) & (4,5)   \\
+\hline
+$K_A$  & 459.88 & 447.12 & 469.32 & 494.40 & 565.20  \\
+\hline
+$K_S$  & 458.09 & 459.38 & 466.44 & 502.08 & 570.70  \\
+\hline
+$T_A$  & 114.97 &  55.89 &  39.11 &  30.90 &  28.26  \\
+\hline
+$T_S$  & 144.45 &  71.38 &  46.58 &  36.92 &  33.69  \\
+\hline
+$u_A$  &   3.61 &   7.43 &  10.61 &  13.43 &  14.69  \\
+\hline
+$u_S$  &   2.87 &   5.82 &   8.91 &  11.24 &  12.32  \\
+\hline
+$u$    &   1.26 &   1.28 &   1.19 &   1.19 &   1.19  \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+The next series of the experiments consists in the minimization of the multidimensional function \cite{5_LucidiPiccioni}
+\begin{equation}
+\label{eq:5_33}
+\varphi(y)=\frac{\pi}{N}\left\{10\sin^2(\pi y_1)+(y_N-1)^2+\sum_{i=1}^{N-1}{[(y_i-1)^2)(1+10\sin^2(\pi y_{i+1})]}    \right\}
+\end{equation}
+where $-2\leq y_i\leq 4,\;1\leq i\leq N$, for the dimensions $N=$ 3, 4, and 5.
+The precision $\epsilon=0.12$ and the reliability parameter $r = 2$ were used in all the experiments.
+At the initial search steps the first four trials were conducted at the points {-1.2, -0.5, 0.0, 2.2} in all one-dimensional subproblems (\ref{eq:5_21}).
+The results of the computations for the parallel algorithms are presented in Table \ref{tab:5_2}.
+\begin{table}
+\caption{Comparison of the parallel algorithms for the functions (\ref{eq:5_33}) of various dimensions}
+\label {tab:5_2}
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+\multirow{8}{*}{$N=3$} & $q$ & (1,1,1) & (2,2,2) & (3,3,3) & (4,4,4) \\
+\cline{2-6}
+		& $K_A$ &  7 791  &  7 056  &  7 884  &  10 368   \\
+\cline{2-6}
+		& $K_S$ &  7 791  &  6 772  &  7 548  &  10 836   \\
+\cline{2-6}
+		& $T_A$ &  7 791  &   882  &   292  &    162   \\	
+\cline{2-6}
+		& $T_S$ &  7 791  &  1 013  &   390  &    216   \\		
+\cline{2-6}
+		& $u_A$ &  1.00  &  8.83  & 26.68  &  48.09   \\
+\cline{2-6}
+		& $u_S$ &  1.00  &  7.69  & 19.98  &  36.07   \\
+\cline{2-6}
+		& $u$   &  1.00  &  1.15  &  1.34  &   1.33   \\
+\hline
+\hline
+\multirow{8}{*}{$N=4$} & $q$ & (1,1,1,1) & (2,2,2,2) & (3,3,3,3) & (4,4,4,4) \\  
+\cline{2-6}
+		& $K_A$ &  155 694  &  136 576  &  165 807  &  201 216   \\
+\cline{2-6}
+		& $K_S$ &  155 694  &  130 168  &  152 895  &  232 682   \\
+\cline{2-6}
+		& $T_A$ &  155 694  &  8 536  &   2 047  &   786   \\	
+\cline{2-6}
+		& $T_S$ &  155 694  & 10 658  &  3 062  &    1 296   \\		
+\cline{2-6}
+		& $u_A$ &  1.00  & 18.24  & 76.06  & 198.08   \\
+\cline{2-6}
+		& $u_S$ &  1.00  & 14.61  & 50.85  & 120.13   \\
+\cline{2-6}
+		& $u$   &  1.00  &  1.25  &  1.50  &   1.65   \\
+\hline
+\hline
+\multirow{8}{*}{$N=5$} & $q$ & (1,1,1,1,1) & (2,2,2,2,2) & (3,3,3,3,3) & (4,4,4,4,4) \\  
+\cline{2-6}
+		& $K_A$ &  3111771  &  2635872  &  3258144  &  4786176   \\
+\cline{2-6}
+		& $K_S$ &  3111771  &  2502128  &  3097143  &  4996416   \\
+\cline{2-6}
+		& $T_A$ &  3111771  &  82371  &   13408  &   4674   \\	
+\cline{2-6}
+		& $T_S$ &  3111771  & 112485  &  24119  &    7776   \\		
+\cline{2-6}
+		& $u_A$ &  1.00  & 37.78  & 232.08  & 665.76   \\
+\cline{2-6}
+		& $u_S$ &  1.00  & 27.66  & 129.02  & 400.18   \\
+\cline{2-6}
+		& $u$   &  1.00  &  1.36  &  1.80  &   1.67   \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+In the next series of experiments, the function computation time was assumed to depend on the trial point.
+Problem (\ref{eq:5_33}) was considered for  $N=3$; the precision of the method $\epsilon=0.01$ and the reliability parameter $r=2$ were taken.
+In all one-dimensional subproblems (\ref{eq:5_21}) the first four trials were conducted at the points {-1.2, -0.5, 0.0, 2.2}.
+The two kinds of the dependencies $t(y)$ of the function value computation time per one trial point were used.
+In experiment A
+\begin{displaymath}
+t_1(y)=\left\lfloor 999\frac{\varphi(y)-\varphi_{min}}{\varphi_{min}-\varphi_{max}}\right\rfloor,
+\end{displaymath}
+where $\left\lfloor x\right\rfloor$  is the integer part of $x$, $\varphi_{min}=0$, $\varphi_{max}=185$, and in experiment B
+\begin{displaymath}
+t_2(y)=1000-t_1(y)
+\end{displaymath}
+were used.
+
+Table 5.3 contains the results of the experiments A and B.
+\begin{table}
+\caption{Comparison of parallel algorithms in the case of variable time of trial execution}
+\label {tab:5_3}
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+ & \multicolumn{2}{c|}{Experiment A} & \multicolumn{2}{c|}{Experiment B} \\
+\hline
+$q$ & (2,2,2) & (3,3,3) & (2,2,2) & (3,3,3)    \\
+\hline
+$K_A$ & 106 802 & 117 270 & 105 599 & 111 774 \\
+\hline
+$K_S$ & 101 510 & 103 671 & 101 510 & 103 671 \\
+\hline
+$T_A$ & 738 361 & 221 739 & 11 900 017 & 3 629 895 \\
+\hline
+$T_S$ & 1 527 229 & 791 559 & 16 943 876 & 5 769 930 \\
+\hline
+$u$   & 2.07  &  3.57  &  1.42  &  1.59  \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+It results from Table \ref{tab:5_3} that both the methods make a small number of redundant trials only, and the asynchronous algorithm is faster than the synchronous one.
+
+Now let us consider a parallel characteristical algorithm.
+It is possible construct its generalization onto the synchronous and asynchronous cases like it has been done in Chapter 3 and compare the speed-ups of these versions as a measure of parallelization efficiency.
+This approach allows estimating the potential of various parallelization schemes of the algorithm decision rules.
+In the framework of this approach, the experiments were carried out for core information global search algorithm (\ref{eq:2_36})--(\ref{eq:2_38}) for the  class of functions of various dimensionality.
+The average values of $u_S(\pi)$ and $u_A(\pi)$ obtained by the optimization of 100 multiextremal two-dimensional functions selected randomly from the test class (\ref{eq:2-58}) are presented in Fig.
+\ref{fig:5_7}.
+\begin{figure}[t]
+\centering
+%\includegraphics[width=0.8\linewidth]{figures/figure_5_7.pdf}
+\caption{Comparative speed-ups of the synchronous scheme and of the asynchronous one on the two-dimensional test function class}
+\label{fig:5_7}    
+\end{figure}
+The function plots demonstrate the dependencies of the speed up (the ordinate axis) on parallelization vector (\ref{eq:5_32}), the variants of which are shown on the abscissa axis.
+The upper plot corresponds to the «ideal» speed up.
+The middle and the lower plots reflect the speed-ups of the asynchronous scheme and of the synchronous one, respectively.
+
+An analogous graph for the case of the optimization of a multiextremal five-dimensional function
+\begin{displaymath}
+\varphi(y)=\frac{\pi}{5}\left(\sin^2(\pi y_1)+5(y_5-1)^2+\sum_{i=1}^4{[(y_i-1)^2)(1+10\sin^2(\pi y_{i+1})]}    \right)
+\end{displaymath}
+is presented in Fig.
+\ref{fig:5_8}.
+For the one-dimensional optimization the parallel version of the broken lines method (\ref{eq:2_36}), (\ref{eq:2_37}) with the adaptive estimate of Lipschitz constant (\ref{eq:2_39}), (\ref{eq:2_40}) was used.
+\begin{figure}[t]
+\centering
+%\includegraphics[width=0.8\linewidth]{figures/figure_5_8.pdf}
+\caption{Speed-ups of the synchronous and asynchronous broken lines method}
+\label{fig:5_8}    
+\end{figure}
+
+As in the previous case, the upper plot corresponds to the «ideal» speed up. The middle and lower plots reflect the speed-up of the asynchronous scheme and of the synchronous one, respectively. 
+\section{Adaptive nested scheme of dimensionality reduction}
+\label{sec:5_4}
+\subsection {General description}
+\label{subsec:5_4_1}
+The univariate subproblems of the family (\ref{eq:5_21}) are generated recursively and it is possible to describe their connections using a hierarchical structure “tree” in which  the root is the subproblem (\ref{eq:5_19}) and the leaves are the subproblems (\ref{eq:5_20}).
+An example of the full subproblems’  tree (obtained after completion of solving all the family (\ref{eq:5_21})) for the dimension $N=3$  is shown in Fig.
+\ref{fig:5_9}.
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\linewidth]{figures/figure_5_9.pdf}
+\caption{Tree of subproblems in the nested optimization scheme}
+\label{fig:5_9}    
+\end{figure}
+
+In the classical implementation of the nested scheme there are some important features.
+Namely, execution of any trial in a subproblem at the level $i<N$ generates a sub-tree of the general tree (the execution of a trial in a subproblem at the level $i=N$ consists in computation of the objective function $\varphi(y)$ from (\ref{eq:5_1})).
+For example, in the root subproblem (\ref{eq:5_19})  the trial at the point $y_1^s$  requires solving univariate subproblems of the sub-tree marked in Fig.\ref{fig:5_9} with the background as the dashed lines.
+Moreover, until the subproblems of this sub-tree will be completely solved no new trial can be initiated in the root subtask.
+As a consequence, while solving the subproblems belonging to the dashed sub-tree the subproblems corresponding to the other sub-trees either have been solved already (they are marked in Fig \ref{fig:5_9} with grey) or will be solved later (marked with no background).
+
+As it follows from the presented structure the subproblems at the same level are solved independently and the search information obtained by a subproblem is not used in the other subproblems at this level (and, of course, in all sub-trees generated them).
+Such loss of information leads to increasing the number of trials and slows up the optimization process.
+
+To overcome this drawback in the paper \cite{5_GerGriGer} a new version of the nested optimization scheme in which the other way of solving the univariate subproblems has been proposed.
+In this version called adaptive nested optimization scheme instead of the strictly subordinated order of solving the univariate subproblems the principle of simultaneous consideration of all subproblems arising dynamically is used.
+The adaptive scheme does not require to solve each  one-dimensional subproblem  up to the end and all the generated subproblems participate in the multidimensional optimization at any iteration of its realization.
+The main principle of such realization is close to the idea of characteristical algorithms.
+Namely, to each subproblem a numerical value called its characteristic is assigned and the multidimensional iteration consists in the choice of the subproblem with maximal characteristic and performing a new trial in this subtask.
+If the characteristics are assigned in such a way that they take into account the values of functions minimized in subproblems following the principle “low values influence increasing the characteristic” then the subproblems with lower values will be chosen more often than the subproblems with the higher ones and computational resources will not be spent for solving the latters unlike the classical nested scheme where these subproblems have to be solved completely.
+
+The detailed algorithmic implementation of the adaptive nested scheme can be found in \cite{5_GerGriGer}. Here we give a description of this scheme in a general form.
+
+\textbf{Start stage.}
+
+Create in some way an initial set $S$  of subproblems (\ref{eq:5_21}) including a few subproblems at each level of recursion.
+\begin{description} [\textbf{Step G1.}]
+\item [\textbf{Step G1.}] {Juxtapose to each subproblem from the set $S$ a numerical value called characteristic of this problem.}
+\item [\textbf{Step G2.}] {Choose the subproblem with the greatest characteristic (the currently best subproblem).}
+\item [\textbf{Step G3.}]{Compute a new trial point in the chosen subproblem.}
+\item [\textbf{Step G4.}]{Build the sub-tree of subtasks generated by this trial point in accordance with the nested scheme  (\ref{eq:5_18})--(\ref{eq:5_21}) and add new subproblems to the set $S$.}
+\item [\textbf{Step G5.}]{If the termination  criterion is fulfilled in the root subproblem (\ref{eq:5_19}) stop the search and take as a result of optimization the least computed value of the objective function $\varphi(y)$ and its coordinates as the global minimizer.
+Otherwise, go to \textbf{Step G1}.}
+\end{description}
+
+The crucial question arising in the general scheme is how to choose characteristics of the subproblems.
+A possible way can be suggested if for solving the subproblems (\ref{eq:5_21}) a one-dimensional characteristical algorithm is used.
+In this case as the characteristic of the subtask one can take the best interval characteristic (\ref{eq:2_31}) of this method.
+Moreover, as a stopping criterion of the multidimensional search the termination criterion (\ref{eq:2_21}) in the root problem (\ref{eq:5_19}) can be taken, i.e., the multidimensional optimization is stopped when the interval with the maximal characteristic in the subproblem (\ref{eq:5_19}) becomes less than predefined accuracy.
+
+For realization of this idea the core information  algorithm AGS (\ref{eq:2_37})--(\ref{eq:2_39}) can be taken.
+Theorem ~\ref{theor:5_2} gives the theoretical substantiation of convergence to global minimum for the adaptive nested scheme combined with AGS for solving internal univariate subproblems (\ref{eq:5_21}).
+
+Before formulating the theorem let us introduce a few notations.
+First of all, it is necessary to pay attention that during realization of the adaptive scheme the values of objective functions   in subproblems (\ref{eq:5_21}) for $1\leq i\leq N-1$   are evaluated approximately, as for obtaining the exact value $\varphi^i(u_{i-1},\tilde{y_i})$ at a
+point  $\tilde{y_i}$  it is required to minimize exactly the function $\varphi^{i+1}(u_{i-1},\tilde{y_i},y_{i+1})$  
+at the next level of recursion. So, instead of exact values  $\varphi^i(u_{i-1},y_i^q),\:1\leq q\leq k_i$,   at trial 
+points $y_i^1,\ldots,y_i^{k+i}$  we have the values  $\psi^i(u_{i-1},y_i^q),\:1\leq q\leq k_i$,  such that 
+\begin{equation}
+\label{eq:5_34}
+\left|\varphi^i(u_{i-1},y_i^q)-\psi^i(u_{i-1},y_i^q)\right|\leq \gamma_q^i,1\leq q\leq k_i,
+\end{equation}
+where $\gamma_q^i$ can be interpreted as errors of evaluations for the exact values $\varphi^i(u_{i-1},y_i^q)$, $1\leq q\leq k_i$.
+As a consequence, after termination of the multidimensional optimization instead of global minimum $\varphi^*$  of the function (\ref{eq:5_1})  there is an approximate estimate $\psi^*$ of it.
+In this regard the question on the proximity of values $\varphi^*$ and  $\psi^*$ is very important.
+The answer gives the following theorem (see \cite{5_GerGriGer}).
+\begin{theorem}
+\label{theor:5_2}
+Let the problem (\ref{eq:5_1})--(\ref{eq:5_3}) be considered without functional constraints (\ref{eq:5_2}), i.e., the admissible domain $Q$ coincides with the hyperparallelepiped $D$, and in the adaptive scheme  the information algorithm AGS (\ref{eq:2_37})--(\ref{eq:2_39}) be applied for one-dimensional optimization (\ref{eq:5_21}).
+If
+\begin{enumerate}
+\item {the objective function $\varphi(y)$  of the problem (\ref{eq:5_1}) satisfies in the domain $D$  the Lipschitz condition (\ref{eq:5_29}) with a finite constant $L>0$;}
+\item {during solving the subproblems (\ref{eq:5_21}) the parameter $m$ from (\ref{eq:2_39}) satisfies the inequality $m>2L$;}
+\item {while solving the root problem (\ref{eq:5_19}) the termination criterion (\ref{eq:2_21}) is used with $\epsilon=0$;}
+\item {for any function $\varphi^i(u_{i-1},y_i)$  of the family (\ref{eq:2_21}) the conditions 
+\begin{equation}
+\label{eq:5_35}
+\gamma_{\tau-1}^i+\gamma_\tau^i\leq 2\varphi(y^*)+L(y_i^\tau-y_i^{\tau-1})--(\varphi(u_{i-1},y_i^{\tau-1}+\varphi(u_{i-1},y_i^\tau)
+\end{equation}
+hold for the computational errors $\gamma_{\tau-1}^i,\gamma_\tau^i$  of trials at the adjacent points $ y_i^{\tau-1},y_i^\tau$ such that
+\begin{displaymath}
+y_i^{\tau-1}\leq y_i^*\leq y_i^\tau,
+\end{displaymath}
+where $y^*=(y_1^*,\ldots,y_N^*)$  is the global minimizer of the problem (\ref{eq:5_1})--(\ref{eq:5_3}),}
+\end{enumerate}
+then the sequence of multidimensional trials generated by the adaptive scheme converges to the global minimizer $y^*$ .
+\end{theorem}
+
+This theorem provides sufficient conditions of convergence to global minimum for the adaptive scheme in the combination with the method AGS.
+The proof of the theorem can be found in \cite{5_GerGriGer}.
+Analogous theorem for the adaptive nested scheme with combination of the univariate broken lines method \cite{5_Piyavskij} has been given in \cite{5_GriIsrSergAMC}.
+\subsection {Numerical experiments}
+\label{subsec:5_4_2}
+To demonstrate the efficiency of the adaptive nested scheme let us consider a function from the 2-dimensional test class (\ref{eq:3-58}) (see \cite{5_GrishaginOperChar, 5_StrSergMon2000}).
+In Fig.~\ref{fig:5_10} the level curves of this function and distribution of trials (marked by points) are presented.
+The left panel corresponds to the adaptive scheme and the right panel to the classical one.
+For optimization of the function these schemes are used in combination with the method AGS for solving the univariate subproblems (\ref{eq:5_21}).
+In both the cases the method parameter $r=3$  and the search accuracy $\epsilon=0.01$ in termination criterion (\ref{eq:2_21}) were used.
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\linewidth]{figures/figure_5_10.png}
+\caption{Trials distribution of the adaptive scheme (the left panel) and of the classical one (the right panel).}
+\label{fig:5_10}    
+\end{figure}
+
+Both the schemes have provided the required accuracy of the problem solution.
+However, the adaptive scheme has spent significantly less trials, namely, the classical scheme evaluated the objective function 1139 times while the adaptive version carried out 257 trials only.
+
+Another experiment was executed \cite{5_GriIsrSergAMC} on the set of 100 three-dimensional functions belonging to the well-known test class GKLS \cite{5_GavianoKvasovLeraSergeyev}.
+The GKLS parameters of the test functions were as follows:
+\begin{itemize}
+\item {10 local minima;}
+\item {standard function value $-1.0$ at the  global minimizer;}
+\item{distance 0.9 from the global minimizer to the vertex of the paraboloid;}
+\item{radius 0.12 of the attraction region of the global minimizer.}
+\end{itemize}
+
+The following algorithms were compared by means of the method of operational characteristics – the adaptive nested scheme (AGS-A) and the classical nested scheme (AGS-C) with the univariate method AGS inside of both the schemes, and the very popular method DIRECT \cite{5_Jones, 5_JonesPerttunenStuckman} as an example of global optimization methods of different nature.
+Operational characteristics (see Subsection ~\ref{subsec:2_2_3} of Chapter 2) of the methods compared are presented in Fig. ~\ref{fig:5_11}.
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\linewidth]{figures/figure_5_11.pdf}
+\caption{Operational characteristics of the nested optimization schemes and the method DIRECT.}
+\label{fig:5_11}    
+\end{figure}
+
+As it follows from the presented examples and it is confirmed in other researches (see, for instance, \cite{5_GerGriGer, 5_GriIsrSergAIP, 5_GriIsrSergAMC}) the adaptive nested scheme improves significantly the results of optimization compared with its classical prototype.
+
+\subsection{Parallel adaptive scheme of nested optimization}
+\label{subsec:5_4_3}
+As well as the classical nested scheme, the adaptive one can be parallelized.
+The idea of parallelizing is similar to the approach applied in the parallel characteristical algorithms.
+Namely, when in a multiprocessor environment there are $p>1$ free processors, in the decision rule of the general stage for the choice of the subproblems for further solving one can take $p$ subproblems with the best characteristics.
+For more details, let us reformulate for the parallel case the general description of the sequential adaptive scheme modifying Steps G2-G4.
+
+\paragraph{\textbf{General iteration of synchronous adaptive scheme}}
+\begin{description} [\textbf{Step G1.}]
+\item [\textbf{Step G1.}] {Juxtapose to each subproblem from the set $S$  a numerical value called characteristic of this problem.}
+\item [\textbf{Step G2.}] {Choose $p$ subproblems with the greatest characteristics.}
+\item [\textbf{Step G3.}]{Compute new trial points in the chosen subproblems and assign to each point one processor from the pool of free processors.}
+\item [\textbf{Step G4.}]{Build in parallel the sub-trees of subtasks generated by these trial points in accordance with the nested scheme (\ref{eq:5_18})--(\ref{eq:5_21}) and add new subproblems of these sub-trees to the set $S$.}
+\item [\textbf{Step G5.}]{If the termination  criterion is fulfilled in the root subproblem (\ref{eq:5_19}) stop the search and take as a result of optimization the least computed value of the objective function $\varphi(y)$ and its coordinates as the global minimizer.
+Otherwise, go to  \textbf{Step G1}.}
+\end{description}
+
+In this description, the computational parallel procedure is implicitly supposed to be synchronous since a new iteration G1-G5 can start after completion of  building all the sub-trees at Step G4 only.
+Obviously, the times of building those can differ significantly.
+As a result, the processors completing computations wait until the slowest processor finishes and the new iteration of the general computational procedure can go on thereafter only.
+To avoid standing idle one can to use an asynchronous version of the parallel adaptive scheme presented below.
+
+\paragraph{\textbf{General iteration of asynchronous adaptive scheme}}
+\begin{description} [\textbf{Step G1.}]
+\item [\textbf{Step G1.}] {If there are $1\leq\pi\leq p$ free processors go to \textbf{Step G2}.
+Otherwise, if there are no accessible processors, wait until $\pi,\:1\leq\pi\leq p$ processors have completed computations.
+After that add to the set $S$ new subproblems generated by processors having finished and go to the next \textbf{Step G2}.}
+\item [\textbf{Step G2.}] {If the termination criterion is fulfilled in the root subproblem (\ref{eq:5_19}) stop the search and take as a result of optimization the least computed value of the objective function $\varphi(y)$ and its coordinates as the global minimizer.
+Otherwise, go to \textbf{Step G3}.}
+\item [\textbf{Step G3.}]{Juxtapose to each subproblem from the set $S$ a numerical value called characteristic of this problem and choose $\pi$  subproblems with the greatest characteristic (the currently best subproblems).}
+\item [\textbf{Step G4.}]{Compute new trial points in the chosen subproblems and assign to each point one processor from the pool of free processors.}
+\item [\textbf{Step G5.}]{Begin solving subproblems generated by these trial points in accordance with the nested scheme  (\ref{eq:5_18})--(\ref{eq:5_21}) and go to \textbf{Step G1}.}
+\end{description}
+
+The description of general computational procedure for both the synchronous and asynchronous adaptive schemes is just a “skeleton” which requires many important details for a real implementation.
+For example, how to realize the stage of initialization, how to keep in memory the set of subproblems $S$, or how to organize the information exchange between processors?
+It should be noted as well that in adaptive scheme the subproblems of a sub-tree do not required to be solved up to the end, i.e., up to fulfillment of termination criterion in univariate algorithm solving them as it is done in the classical nested scheme.
+In this situation the following question arises: how to stop the one-dimensional search during solving subproblems (\ref{eq:5_21})?
+Moreover, the implementation of the adaptive scheme depends in many ways on the univariate methods use inside it.
+These and other peculiarities of possible implementation indicate a considerable variability of the adaptive scheme and require further investigations of increasing efficiency by means of new combinations of the general scheme with different univariate methods.
+
+In conclusion some comparative results for the sequential and asynchronous versions of the adaptive scheme are presented.
+For the experiment the test class GKLS \cite{5_GavianoKvasovLeraSergeyev} was taken and test sets of functions for dimensions 7 and 8 were used.
+For univariate optimization inside the multidimensional search the information statistical method AGS was applied in all the cases.
+Computations were executed on a cluster consisting of 4 nodes with 4 processors Intel® Xeon® E7-8890 v4 and each processor contained 24 cores.
+Parallelization was realized on the base of MPI in the version Intel® MPI 2017.
+The interconnector Infiniband FDR was used for data exchange between processors.
+
+Fig. ~\ref{fig:5_12} shows the speed-up in time of the asynchronous adaptive scheme as compared to the sequential one.
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.05\linewidth]{figures/figure_5_12.pdf}
+\caption{Speed-up of the asynchronous adaptive scheme.}
+\label{fig:5_12}    
+\end{figure}
+
+Presented results demonstrate the substantial acceleration of the optimization process if to use the parallel adaptive scheme.
+It should be noted that if the parallelism is realized  with  few number of ranks then the efficiency of parallelizing is better for problems of the less dimension, however, the growth of number of MPI ranks leads to increasing the speed-up for  problems with more dimension (for 128 ranks speed-up in 8-dimensional problems is almost 1.5 times more than in 7-dimensional case).
+
+\begin{thebibliography}{99.}
+\bibitem{5_BarkGer2014} Barkalov, K.A., Gergel, V.P.: Multilevel scheme of dimensionality reduction for parallel global search algorithms.
+OPT-i 2014 -- 1st International Conference on Engineering and Applied Sciences Optimization, Proceedings, 2111--2124 (2014)
+\bibitem{5_BarkGerLeb} Barkalov, K., Gergel, V., Lebedev, I.: Solving global optimization problems on GPU cluster  AIP Conference Proceedings \textbf{1738}, 400006 (2016)
+\bibitem{5_BarkLeb} Barkalov, K., Lebedev, I.: Local tuning in multilevel scheme of parallel global optimization.  AIP Conference Proceedings \textbf{1776}, 060006 (2016)
+\bibitem{5_CarrHowe}	Carr, C.R., Howe, C.W.: Quantitative Decision Procedures in Management and Economic: Deterministic Theory and Applications. McGraw-Hill, New York (1964)
+\bibitem{5_Evtushenko}	Evtushenko, Yu.G.: Numerical Optimization Techniques. Translation Series in Mathematics and Engineering. Optimization Software  Inc., Publication Division, New York (1985)
+\bibitem{5_GavianoKvasovLeraSergeyev} Gaviano, M., Kvasov, D.E., Lera, D., Sergeyev, Y.D.: Algorithm 829: Software for generation of classes of test functions with known local and global minima for global optimization.
+ACM Trans.
+Math.
+Software \textbf{29}(4), 469–480 (2003)
+\bibitem{5_GerGriIsr} Gergel, V.,  Grishagin, V., Israfilov, R.: Local tuning in nested scheme of global optimization, Procedia Computer Science \textbf{51}, 865--874 (2015) 
+\bibitem{5_GerGriGer}	Gergel, V., Grishagin, V., Gergel, A.: Adaptive nested optimization scheme for multidimensional global search, J. Glob. Opt. \textbf{66}, 35–51 (2016).
+\bibitem{5_GrishaginOperChar}  Grishagin,V.A.: Operation characteristics of some global search algorithms, Problems of Statistical Optimization \textbf{7},  198--206 (1978). In Russian
+\bibitem{5_GriIsrAIP}	Grishagin, V.A., Israfilov, R.A.: Global search acceleration in the nested optimization scheme. AIP Conf. Proc. \textbf{1738}, 400010 (2016)
+\bibitem{5_GriIsrSergAIP} Grishagin, V., Israfilov, R., Sergeyev, Y.: Comparative efficiency of dimensionality reduction schemes in global optimization. AIP Conf. Proc. \textbf{1776}, 060011 (2016)
+\bibitem{5_GriIsrSergAMC}	Grishagin, V.,  Israfilov, R., Sergeyev, Y.: Convergence conditions and numerical comparison of global optimization methods based on dimensionality reduction schemes.
+Appl.
+Math.
+Comput.
+(2017) doi:10.1016/j.amc.2017.06.036
+\bibitem{5_GriIsrCEUR}Grishagin, V.A., Israfilov, R.A.: Multidimensional Constrained Global Optimization in Domains with Computable Boundaries. CEUR Workshop Proceedings \textbf{1513}, 75--84 (2015)
+\bibitem{5_GrishaginSergeyevStrongin} Grishagin, V.A., Sergeyev, Y.D., Strongin, R.G.: Parallel characteristical algorithms for solving problems of global optimization.
+J.
+Global Optim.
+\textbf{10}(2), 185–206 (1997)
+\bibitem{5_GrishaginStrongin_EnginCybernetics} Grishagin, V.A., Strongin, R.G.: Optimization of multiextremal functions subject to monotonically unimodal constraints.
+Engineering Cybernetics \textbf{22},  117--122 (1984)
+\bibitem{5_Jones}Jones, D.R.: The DIRECT global optimization algorithm. In: C.A. Floudas, P.M. Pardalos
+(eds.) Encyclopedia of Optimization, vol. 1, pp. 431–440. Kluwer Academic Publishers, Dordrecht (2001)
+\bibitem{5_JonesPerttunenStuckman}Jones, D.R., Perttunen, C.D.,   Stuckman, B.E.: Lipschitzian optimization
+without the Lipschitz constant, J. Optim. Theory Appl. 79,  157--181 (1993)
+\bibitem{5_Kushner}	Kushner, H.J.: A new method of locating the maximum point of an arbitrary multipeak curve in the presence of noise.
+Transactions of ASME, Ser.
+D.
+Journal of Basic Engineering \textbf{86}, 97--106 (1964)
+\bibitem{5_Locatelli} Locatelli, M.: Bayesian algorithms for one-dimensional global optimization. Journal of Global Optimization \textbf{1}, 57--76 (1997)
+\bibitem{5_LucidiPiccioni} Lucidi, S., Piccioni, M.:  Random tunneling by means of acceptance-rejection sampling for global optimization. J. Optim. Theory Appl. \textbf{62}(2), 255--277 (1989)
+\bibitem{5_Mockus1980} Mockus, J.: The simple Bayesian algorithm for the multidimensional global optimization.
+In: Archetti, F., and Cugiani, M.
+(Eds.) Numerical Techniques for Stochastic Systems.
+North-Holland, Amsterdam,
+the Netherlands (1980)
+\bibitem{5_Mockus1988}	Mockus, J.: Bayesian Approach to Global Optimization. Kluwer Academic Publishers, Dordrecht (1988)
+\bibitem{5_MockusEddy_et_al} Mockus, J., Eddy, W., Mockus, A., Mockus, L., Reklaitis, G.: Bayesian Heuristic Approach
+to Discrete and Global Optimization. Kluwer Academic Publishers, Dordrecht (1996)
+\bibitem{5_Piyavskij}	Piyavskij, S.A.: An Algorithm for Finding the Absolute Extremum of a Function. USSR Comput. Math. Math. Phys. \textbf{12}(4), 57--67 (1972)
+\bibitem{5_SergeyevLocTun} Sergeyev, Y.D.: An information global optimization algorithm with local tuning. SIAM J. Optim. \textbf{5}(4), 858–870 (1995)
+\bibitem{5_Sergeyev1998} Sergeyev, Y.D.: Global one-dimensional optimization using smooth auxiliary functions. Math. Program. \textbf{81}(1), 127–146 (1998)
+\bibitem{5_SergeyevGrishaginOMS} Sergeyev, Y.D., Grishagin, V.A.: Sequential and parallel algorithms for global optimization, Optim. Methods Softw. \textbf{3},  111--124 (1994)
+\bibitem{5_SergGriJCAA}	Sergeyev, Y.D., Grishagin, V.A.: Parallel Asynchronous Global Search and the Nested Optimization Scheme. J. Comp. Analysis Appl. \textbf{3}, 123--145 (2001)
+\bibitem{5_SergMukhKvasLera} Sergeyev, Y.D., Mukhametzhanov, M.S., Kvasov, D.E., Lera, D.: Derivative-free local tuning and local improvement techniques embedded in the univariate global optimization.
+J.
+Optim.
+Theory Appl.
+171(1), 186--208 (2016)
+\bibitem{5_ShiOlaf}   Shi, L., {\'O}lafsson, S.: Nested Partitions Method for Global Optimization. Operations Research \textbf{48}, 390--407 (2000)
+\bibitem{5_Shubert} Shubert, B.O.: A sequential method seeking the global maximum of a function. SIAM J. Numer. Anal. \textbf{9}(3), 379–388 (1972)
+\bibitem{5_StrMonRus}	Strongin, R.G.: Numerical Methods in Multiextremal Problems (Information-Statistical Algorithms). Nauka, Moscow (1978). In Russian
+\bibitem{5_Str1973} Strongin, R.G.: On the convergence of an algorithm for finding a global extremum. Engineering Cybernetics 11, 549–555 (1973)
+\bibitem{5_StrMarkin} Strongin, R.G., Markin, D.L.: Minimization of multiextremal functions with nonconvex constraints. Cybernetics 22, 486–493 (1986)
+\bibitem{5_StrSergMon2000} Strongin, R.G., Sergeyev, Y.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000) 
+\bibitem{5_SysoyevBarkGerLeb} Sysoyev, A., Barkalov, K., Gergel, V., Lebedev, I.: MPI implementation of dimension reduction multilevel scheme for parallel solving the global optimization problems  CEUR Workshop Proceedings \textbf{1482}, 61--68 (2015)
+\bibitem{5_vanDam}	van Dam, E.R., Husslage, B., Hertog, D. One-dimensional Nested Maximin Designs. J. Glob. Opt. \textbf{46}, 287--306 (2010)
+\bibitem{5_Zilinskas1975} $\check{Z}$ilinskas, A.: One-step Bayesian method for the search of the optimum of one-variable functions. Cybernetics 1, 139--144 (1975)
+\bibitem{5_Zilinskas1981} $\check{Z}$ilinskas, A.: Two algorithms for one-dimensional multimodal minimization. Mat. Operationsforsch. Statist. \textbf{12}, Ser. Optimization, 53--63 (1981)
+\bibitem{5_Zilinskas1985} $\check{Z}$ilinskas, A.: Axiomatic characterization of a global optimization algorithm and investigation of its search strategy.
+Operations Research Letters \textbf{4}, 35--39 (1985)
+\end{thebibliography}
 

### 2017-10-06 2017/ParOptBook/Chapter4.tex
parOptBook: eliminate all hardcoded references in chapter 4

@@
 \begin{equation}
+  \label{eq4:30}
   Q_j=\{x\in[a,b]:g_j(x)\le 0\},\: 1\le j\le m,
@@
 
-\textbf{The constraints’ check order.} he numbering of the constraints and the objective function introduced in the statement of problem \eqref{eq4:problem} and defined by the sequence of integer numbers
+\textbf{The constraints’ check order.} The numbering of the constraints and the objective function introduced in the statement of problem \eqref{eq4:problem} and defined by the sequence of integer numbers
 \begin{equation}
@@
 \begin{equation}
+  \label{eq4:33}
   H(x_i)=\{j_{i1},\dots,j_{im},j_{i,m+1}\},\:0\le i\le k,
@@
 \begin{equation}
+  \label{eq4:34}
   j_{i,m+1}=m+1,\:0\le i\le k.
@@
 
-Since the constraints’ check order while the trial execution at the point $x$ i is determined by the corresponding tuple $H(x_i )$ from (4.33), the termination of the trial at this point after the computation of the values of l problem functions produces the value $z_i=g_\nu(x_i)$ where $\nu=j_{il}$.
-In this connection, let us use not the serial number $l$ of the last function calculated in this point, as it was accepted in \eqref{eq4:trial} but the index of the last function in the basic numeration (4.32) as the
+Since the constraints’ check order while the trial execution at the point $x$ i is determined by the corresponding tuple $H(x_i )$ from \eqref{eq4:33}, the termination of the trial at this point after the computation of the values of l problem functions produces the value $z_i=g_\nu(x_i)$ where $\nu=j_{il}$.
+In this connection, let us use not the serial number $l$ of the last function calculated in this point, as it was accepted in \eqref{eq4:trial} but the index of the last function in the basic numeration \eqref{eq4:basic_h} as the
 value of $\nu(x_i )$ for the point $x_i$ , i.e.,
 \begin{equation}
+  \label{eq4:35}
   \nu(x_i)=j_{il}.
@@
 \emph{Modification 1 (the fixation of the check order)}.
-Each point $x_i$ from \eqref{eq4:search_data} is assigned the corresponding checking order $H(x_i )$ from (4.33), and this order is included into the algorithm database.
-
-\emph{Modification 2 (the determination of the trial points’ indices).} Each point $x_i$ from \eqref{eq4:search_data} is assigned the value $z_i=g_\nu(x_i)$ where according to (4.35), the index $\nu=\nu(x_i)$ is the basic index from (4.32) corresponding to the value of the last function estimated in this point.
+Each point $x_i$ from \eqref{eq4:search_data} is assigned the corresponding checking order $H(x_i )$ from \eqref{eq4:33}, and this order is included into the algorithm database.
+
+\emph{Modification 2 (the determination of the trial points’ indices).} Each point $x_i$ from \eqref{eq4:search_data} is assigned the value $z_i=g_\nu(x_i)$ where according to \eqref{eq4:35}, the index $\nu=\nu(x_i)$ is the basic index from \eqref{eq4:basic_h} corresponding to the value of the last function estimated in this point.
 These indices are used in all computations based on expressions \eqref{fig:4_13}-\eqref{eq4:characteristic}, \eqref{eq4:next_point}.
@@
 \begin{equation}
+  \label{eq4:36}
   \nu(x_{i-1})=j_{i-1,p}=j_{i,q}\not=\nu(x_i)=j_{i-1,u}=j_{i,x},
@@
 
-If for the indices $p,q, u$ and $s$ from (4.36) the conditions
-\begin{equation}
+If for the indices $p,q, u$ and $s$ from \eqref{eq4:36} the conditions
+\begin{equation}
+  \label{eq4:37}
   p<u,q<s,
@@
 
-Otherwise, when the reverse to (4.37) conditions are fulfilled
-\begin{equation}
+Otherwise, when the reverse to \eqref{eq4:37} conditions are fulfilled
+\begin{equation}
+  \label{eq4:38}
   p>u,q>s,
@@
 case, the value $g_\nu(x_i ) is undefined)$.
-In the cases when conditions (4.37) and (4.38) are not fulfilled, the initial order of calculations prescribed by the sixth rule of the initial algorithm remains.
+In the cases when conditions \eqref{eq4:37} and \eqref{eq4:38} are not fulfilled, the initial order of calculations prescribed by the sixth rule of the initial algorithm remains.
 
@@
 \begin{equation}
+  \label{eq4:39}
   H(a)=H(b)=H,
 \end{equation}
-where $H$ from (4.32). Let
+where $H$ from \eqref{eq4:basic_h}. Let
 \begin{displaymath}
@@
 \end{displaymath}
-are the indices of the boundary points of the interval $(x_{t-1},x_t)$ including the current trial point $x^{k+1}$ (according to the eighth rule) determined according to (4.35).
+are the indices of the boundary points of the interval $(x_{t-1},x_t)$ including the current trial point $x^{k+1}$ (according to the eighth rule) determined according to \eqref{eq4:35}.
 Let us introduce a notation $\{s,H\backslash s\}$ for the tuple obtained from $H$ by excluding the index $s$, shifting all indices $1,2,\dots,s-1$ to the right of one position, and the transfer of the index s excluded from $H$ into the first position, i.e.,
@@
 \end{displaymath}
-Note that according to (4.34), the index $m+1$ always occupies the last position.
+Note that according to \eqref{eq4:34}, the index $m+1$ always occupies the last position.
 
@@
  H(x^{k+1})=\{p,H\backslash p\}, (p=q<m+1)\cup(p<q=m+1), \nonumber \\
+ \label{eq4:40}
  H(x^{k+1})=\{q,H\backslash q\}, p=m+1>q\\
@@
 
-\textbf{Note.} While solving the problem \eqref{eq4:problem} rules (4.39) and (4.40) included into the scheme of ACOM assign the index values from (4.35) to all trial points unambiguously.
+\textbf{Note.} While solving the problem \eqref{eq4:problem} rules \eqref{eq4:39} and \eqref{eq4:40} included into the scheme of ACOM assign the index values from \eqref{eq4:35} to all trial points unambiguously.
 Then, any feasible point $x$ always corresponds to the index $\nu(x)=m+1$.
@@
 \begin{enumerate}
-  \item each domain $Q_j ,1\le j\le m+1$ from (4.30) represents a union of a finite number of segments with a positive length;
+  \item each domain $Q_j ,1\le j\le m+1$ from \eqref{eq4:30} represents a union of a finite number of segments with a positive length;
   \item each function $g_j(x): 1\le j\le m+1$ satisfies Lipschitz condition with the corresponding constant $L_j$;
@@
   \begin{equation}
+    \label{eq4:41}
     0<2\epsilon_\nu<L_\nu(\beta-\alpha),
@@
 
-The distribution of the interval lengths $[\alpha,\beta]$ from (4.41) including the solution points $x^*$ corresponding to the problems from the first set is presented in Fig.
+The distribution of the interval lengths $[\alpha,\beta]$ from \eqref{eq4:41} including the solution points $x^*$ corresponding to the problems from the first set is presented in Fig.
 \ref{fig:4_11}.
@@
   \begin{equation}
+    \label{eq4:54}
     \{(x_{t-1},x_t):t=t(q^s)\},\:s=1,2,\dots,
@@
   \begin{gather}
+    \label{eq4:55}
     \overline x= \cap_{s=1}^\infty[x_{t-1},x_t], \\
+    \label{eq4:56}
     \lim_{p\to\infty}\Delta_t=0, \\
+    \label{eq4:57}
     R(t(q^s))>0,\:s=1,2,\dots, \\
+    \label{eq4:58}
     \lim_{p\to\infty}R(t(q^s))=0,
@@
   \label{eq4:4_61}
-  Then, the property (4.56) is a consequence of inequality \eqref{eq4:4_60}.
+  Then, the property \eqref{eq4:56} is a consequence of inequality \eqref{eq4:4_60}.
 
@@
 If $I_{m+1}= \varnothing$, the index $\omega=m+1$.
-If at that $\nu(x_{i -1} )\not = \nu(x_i )$, then according (4.47) and (4.50), $R(i)=2\Delta_i>0$.
+If at that $\nu(x_{i -1} )\not = \nu(x_i )$, then according \eqref{eq4:characteristic_par} and \eqref{eq4:next_point_2_par}, $R(i)=2\Delta_i>0$.
 In the case when $\nu(x_{i -1} )= \nu(x_i )=\omega$, the estimate
@@
 Thus, at any search step $k\ge 1$ there exists an interval having a strictly positive characteristic.
-Hence, taking into account the rules of selection of the interval with the greatest characteristic \eqref{eq4:MaxR_par}, one can conclude that the intervals from system (4.54) should also have the positive characteristics since these ones undergo splitting up by the trial points falling into these intervals.
-Consequently, conditions (4.57) hold.
-
-  It follows from the trial execution rules and condition (4.35) that for $\nu(x_i )\le m$ the values $z_i$ assigned to the points $x_i$ from \eqref{eq4:par_search_data} are positive.
+Hence, taking into account the rules of selection of the interval with the greatest characteristic \eqref{eq4:MaxR_par}, one can conclude that the intervals from system \eqref{eq4:54} should also have the positive characteristics since these ones undergo splitting up by the trial points falling into these intervals.
+Consequently, conditions \eqref{eq4:57} hold.
+
+  It follows from the trial execution rules and condition \eqref{eq4:35} that for $\nu(x_i )\le m$ the values $z_i$ assigned to the points $x_i$ from \eqref{eq4:par_search_data} are positive.
 Therefore, according to \eqref{eq4:z_const_par},
@@
   \end{displaymath}
-    Now from the rules \eqref{eq4:characteristic_par} for the characteristics and the limit (4.56), taking into account the condition \eqref{eq4:lip_condition} of Lipschitzness of the functions and inequality (4.57), one can derive out the truth of statement (4.58).
+    Now from the rules \eqref{eq4:characteristic_par} for the characteristics and the limit \eqref{eq4:56}, taking into account the condition \eqref{eq4:lip_condition} of Lipschitzness of the functions and inequality \eqref{eq4:57}, one can derive out the truth of statement \eqref{eq4:58}.
 \end{proof}
@@
   Let us assume the opposite and consider the interval $[x_{t-1},x_t]$ including the problem solution $x^*$ from \eqref{eq4:problem} at the step $k$, i.e., $t=t(k)$.
-Since we have assumed the equality \eqref{eq4:62} not to take place and, consequently, the sequence $\{x^k\}$ does not include the points from the feasible domain $Q$, the interval $[\alpha,\beta]$ from (4.52) should satisfy the conditions
+Since we have assumed the equality \eqref{eq4:62} not to take place and, consequently, the sequence $\{x^k\}$ does not include the points from the feasible domain $Q$, the interval $[\alpha,\beta]$ from \eqref{eq4:reserves_par} should satisfy the conditions
   \begin{equation}
@@
 
-  From (4.58), \eqref{eq4:67}, and from the rule (4.48) determining the choice of the interval for the next trial point, the inevitability of carrying out a trial in the interval $[x_{t-1},x_t]$ from \eqref{eq4:63} follows.
+  From \eqref{eq4:58}, \eqref{eq4:67}, and from the rule \eqref{eq4:MaxR_par} determining the choice of the interval for the next trial point, the inevitability of carrying out a trial in the interval $[x_{t-1},x_t]$ from \eqref{eq4:63} follows.
 Consequently, the interval $[x_{t-1},x_t]$ will be divided into two subintervals.
@@
   Since the coincidence $x^*$ with one of the trial points means \eqref{eq4:62} to hold, then it remains to consider the case when $x^*\in(x_{t- 1} ,x_t ),\: t=t(k)$ for any $k\ge 1$.
-Here, because of the first condition of Theorem \ref{th4:convergence_par} and the definition (4.31) either $[x_{t-1},x^*]\in Q$ or $[x^* ,x_t]\in Q$ if the number $k$ is sufficiently large.
+Here, because of the first condition of Theorem \ref{th4:convergence_par} and the definition \eqref{eq4:Q_union} either $[x_{t-1},x^*]\in Q$ or $[x^* ,x_t]\in Q$ if the number $k$ is sufficiently large.
 In this case, either $\nu(x_{t-1})=m+1$ or $\nu (x_t )=m+1.$
@@
 \begin{proof}
-  Let us assume the opposite. According to (4.55) and (4.56),
+  Let us assume the opposite. According to \eqref{eq4:55} and \eqref{eq4:56},
   \begin{displaymath}
@@
   \end{displaymath}
-  where q s is from (4.57) Therefore, it follows from the assumption $x\in Q$ that there exists such large number $S>1$ and such small subinterval
+  where q s is from \eqref{eq4:57} Therefore, it follows from the assumption $x\in Q$ that there exists such large number $S>1$ and such small subinterval
   \begin{equation}
@@
 
-  Let us assign to $Q_j, 1\le j\le m$, from (4.30) their open complements $D_j=[a,b]\setminus Q_j$ consisting of finite number of the open intervals with the positive length (as a consequence of the first condition of Theorem \ref{th4:convergence_par}).
+  Let us assign to $Q_j, 1\le j\le m$, from \eqref{eq4:30} their open complements $D_j=[a,b]\setminus Q_j$ consisting of finite number of the open intervals with the positive length (as a consequence of the first condition of Theorem \ref{th4:convergence_par}).
 For these complements
@@
 
-  According to the trial execution rules and to conditions (4.35) and \eqref{eq4:72}, for any interval from \eqref{eq4:70} the relations
+  According to the trial execution rules and to conditions \eqref{eq4:35} and \eqref{eq4:72}, for any interval from \eqref{eq4:70} the relations
   \begin{displaymath}
@@
   \end{displaymath}
-  hold where $1\le \omega_1,\: \omega_2\le m$. Moreover, according to the rules (4.39) and (4.40) defining the constraints’ check order at trial points,
+  hold where $1\le \omega_1,\: \omega_2\le m$. Moreover, according to the rules \eqref{eq4:39} and \eqref{eq4:40} defining the constraints’ check order at trial points,
   \begin{displaymath}
@@
 
-  Assume $s>1$ to be so large that for $s>S$ the inequality $q^s\ge h$ where $h$ is from \eqref{eq4:65} holds. Additionally, according to (4.46),
+  Assume $s>1$ to be so large that for $s>S$ the inequality $q^s\ge h$ where $h$ is from \eqref{eq4:65} holds. Additionally, according to \eqref{eq4:z_const_par},
   \begin{equation}
@@
   \end{equation}
-  Then, from (4.47), (4.56), \eqref{eq4:71}, and \eqref{eq4:73} we get the estimate
+  Then, from \eqref{eq4:characteristic_par}, \eqref{eq4:56}, \eqref{eq4:71}, and \eqref{eq4:73} we get the estimate
   \begin{equation}
@@
   where $\nu$ is an index from the set $\{\omega_1,\omega_2\}$.
-  Now, from (4.46), \eqref{eq4:74}, and (4.57) one can conclude that if the point $x$ is not a feasible one, it cannot be a limit point of the sequence $\{x^k\}$, i.e., for any limit point $x$ the condition \eqref{eq4:68} is true.
+  Now, from \eqref{eq4:z_const_par}, \eqref{eq4:74}, and \eqref{eq4:57} one can conclude that if the point $x$ is not a feasible one, it cannot be a limit point of the sequence $\{x^k\}$, i.e., for any limit point $x$ the condition \eqref{eq4:68} is true.
 \end{proof}
@@
 
-  From (4.47) taking into account \eqref{eq4:75} and the inequality $z_{m + 1}\ge \varphi(x)$ resulting from (4.46), one can derive the estimate
+  From \eqref{eq4:characteristic_par} taking into account \eqref{eq4:75} and the inequality $z_{m + 1}\ge \varphi(x)$ resulting from \eqref{eq4:z_const_par}, one can derive the estimate
   \begin{equation}
@@
 Let us demonstrate that any limit point $\overline x$ of the sequence $\{x^k\}$ generated by ACOM is a solution of the problem \eqref{eq4:problem} under the conditions of the theorem.
-Since the point $x^*$ from \eqref{eq4:problem} is a limit one, then, according to (4.46),
+Since the point $x^*$ from \eqref{eq4:problem} is a limit one, then, according to \eqref{eq4:z_const_par},
   \begin{equation}
@@
   Let $x\not=x^*$ is another limit point of the sequence $\{x^k\}$ being a feasible one according to \eqref{eq4:68}.
-Then, for the characteristic of the interval $[x_{j-1},x_j]$ including the point $\overline x$ at the step $k$, because of Lipschitz condition \eqref{eq4:lip_condition}, (4.45), (4.47), (4.56), and \eqref{eq4:78} the relation
+Then, for the characteristic of the interval $[x_{j-1},x_j]$ including the point $\overline x$ at the step $k$, because of Lipschitz condition \eqref{eq4:lip_condition}, \eqref{eq4:lip_est_par}, \eqref{eq4:characteristic_par}, \eqref{eq4:56}, and \eqref{eq4:78} the relation
   \begin{displaymath}
@@
 
-  If $\varphi(x)>\varphi(x^*)$ the characteristic of the interval $[x_{j-1},x_j]$ including the point $\overline x$ will become negative starting from some step that contradicts to property (4.57), which should be fulfilled for any limit point.
+  If $\varphi(x)>\varphi(x^*)$ the characteristic of the interval $[x_{j-1},x_j]$ including the point $\overline x$ will become negative starting from some step that contradicts to property \eqref{eq4:57}, which should be fulfilled for any limit point.
 Consequently, the equality $\varphi(x)=\varphi(x^*)$  is true.

### 2017-10-07 2017/ParOptBook/Chapter1.tex
Chapter -> paper

@@
-%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% sample root file for your "contribution" to a contributed volume
+%%
+%% Use this file as a template for your own input.
+%%
+%%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% sample root file for your "contribution" to a contributed volume
-%
-% Use this file as a template for your own input.
-%
-%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %
@@
 
-\abstract{This chapter is devoted to the basic statements and general definitions concerning the global optimization problems and algorithms.
+\abstract{This paper is devoted to the basic statements and general definitions concerning the global optimization problems and algorithms.
 At the beginning, relevant notions are introduced and different statements of global optimization problems are described.
@@
 \begin{equation}
-\label{eq:1}
+\label{eq:1_1_1}
 \inf_{y\in Q}\varphi(y) \equiv\inf\{\varphi(y):y \in Q\}
@@
 \begin{equation}
-\label{eq:2}
+\label{eq:1_1_2}
 \varphi(y^*)=\inf\{\varphi(y):y \in Q\}
@@
 \begin{equation}
-\label{eq:3}
+\label{eq:1_1_3}
 \min_{y\in Q}\varphi(y) \equiv\min\{\varphi(y):y \in Q\}.
@@
 \end{definition}
-Let us denote the set of all points $y\in Q$  satisfying (\ref{eq:2}) as
+Let us denote the set of all points $y\in Q$  satisfying (\ref{eq:1_1_2}) as
 \begin{equation}
-\label{eq:4}
+\label{eq:1_1_4}
 Q^*=Arg\min_{y\in Q}\varphi(y) \equiv Arg\min\{\varphi(y):y \in Q\}.
@@
 \begin{equation}
-\label{eq:5}
+\label{eq:1_1_5}
 \varphi^*=\inf\{\varphi(y):y \in Q\}.
@@
 
-\textbf{Problem Statement B.} Find the infimum from (\ref{eq:5}) and if the set of the global minimum points from (\ref{eq:2}) is not empty, find at least one point $y^*=Q^*$.
+\textbf{Problem Statement B.} Find the infimum from (\ref{eq:1_1_5}) and if the set of the global minimum points from (\ref{eq:1_1_2}) is not empty, find at least one point $y^*=Q^*$.
 
-\textbf{Problem Statement C.} Find the infimum from (\ref{eq:5})  and find all global minimum points (or make sure that the set $Q^*$ is empty).
+\textbf{Problem Statement C.} Find the infimum from (\ref{eq:1_1_5})  and all global minimum points (or make sure that the set $Q^*$ is empty).
 
@@
 
-The problem statements A—D are related to the finding of the global minimum, but sometimes the following problem statement is used even for the multiextremal functions:
+The problem statements A--D are related to the finding of the global minimum, but sometimes the following problem statement is used even for the multiextremal functions:
 
@@
 
-The optimization problem statements A–E are the most commonly used (see, for example, classical monographs \cite{1_FloudasPardalos, 1_Himmelblau, 1_HorstPardalos, 1_HorstTuy, 1_McCormick, 1_Mockus, 1_Pinter,  1_SergKvasMonogr, 1_StrSergMon2000, 1_ZhigZhil}), although other variants are possible as well, for example, those given in \cite{1_Himmelblau}.
+The optimization problem statements A--E are the most commonly used (see, for example, classical monographs \cite{1_FloudasPardalos, 1_Himmelblau, 1_HorstPardalos, 1_HorstTuy, 1_McCormick, 1_Mockus, 1_Pinter,  1_SergKvasMonogr, 1_StrSergMon2000, 1_ZhigZhil}), although other variants are possible as well, for example, those given in \cite{1_Himmelblau}.
 The desired extremal characteristics defined by the problem statement $Z\in\{A, B, C, D, E\}$ are referred to as $Z$-solution of the optimization problem.
@@
 \begin{equation}
-\label{eq:6}
+\label{eq:1_6}
 \varphi(y)\rightarrow\inf, y \in Q.
@@
 
-As  the supremum of the function $\varphi(y)$ in the domain $Q$ is connected with the infimum of the function $-\varphi(y)$ according to  the expression
+As  the supremum of a function $\varphi(y)$ in a domain $Q$ is connected with the infimum of the function $-\varphi(y)$ according to  the expression
 \begin{displaymath}
@@
 the problem of finding the extremal characteristics related to the maximum of the function $\varphi(y)$ (the maximization problem), can be reduced to the problem of the minimization of the function $-\varphi(y)$.
-Therefore, hereinafter the optimization problem will be considered in the form (\ref{eq:6}).
-The function $-\varphi(y)$ in (\ref{eq:6}) will be called \textit{the objective function, the minimized function, or the optimized function},  the domain $Q$ will be called \textit{the feasible (admissible) domain}, and the elements of the set $Q$ will be called \textit{the feasible points}.
+Therefore, hereinafter the optimization problem will be considered in the form (\ref{eq:1_6}).
+The function $\varphi(y)$ in (\ref{eq:1_6}) will be called \textit{the objective function, the minimized function, or the optimized function},  the domain $Q$ will be called \textit{the feasible (admissible) domain}, and the elements of the set $Q$ will be called \textit{the feasible points}.
 
-Let us formulate the problem (\ref{eq:6}), for which the set of the global minimum points $Q^*$ is known to be nonempty a priori (the sufficient conditions for the non-emptiness of $Q^*$ are given, for example, by Weierstrass theorem), as
+Let us denote the problem (\ref{eq:1_6}), for which the set of the global minimum points $Q^*$ is known to be nonempty a priori (the sufficient conditions for the non-emptiness of $Q^*$ are given, for example, by Weierstrass theorem), as
 \begin{equation}
-\label{eq:7}
+\label{eq:1_7}
 \varphi(y)\rightarrow\min, y \in Q.
@@
 
-Note that the mathematical programming problem is often formulated just in this form, although, of course, the problem (\ref{eq:6}) is the more general one.
-In the present monograph, the existence of a global minimizer is always presumed. Let us reformulate the problem (\ref{eq:8}) in the form of a non-linear programming problem
+Note that the mathematical programming problem is often formulated just in this form, although, of course, the problem (\ref{eq:1_6}) is the more general one.
+Let us reformulate the problem (\ref{eq:1_7}) in the form of a non-linear programming problem
 \begin{equation}
-\label{eq:8}
+\label{eq:1_8}
 \varphi(y)\rightarrow\min, y \in Q\subseteq R^N,
@@
 \begin{equation}
-\label{eq:9}
+\label{eq:1_9}
 Q=\{y \in D:g_j(y)\leq 0,1\leq j\leq m \},
@@
 \begin{equation}
-\label{eq:10}
+\label{eq:1_10}
 D=\{y \in R^N:y_i\in [a_i,b_i],1\leq i\leq N \},
 \end{equation}
-i.e., the problem of finding the extremal values (in terms of the Problem statements A-D) of the objective (minimized) function $\varphi(y)$  over the domain $Q$ is defined by the \textit{coordinate} (\ref{eq:10}) and \textit{functional} (\ref{eq:9}) \textit{constraints} on the selection of feasible points (or vectors) $y=(y_1,\ldots ,y_N)$ .
-In this model, the constants $a_i,b_i, 1\leq i\leq N$,  determine the limits for the variation of the problem parameters (the components of the vector $y$), and are finite or (when the corresponding upper or/and lower limits are absent) are assumed to be equal to $a_i=-\infty $  and/or $b_i=+\infty $ .
+i.e., the problem of finding the extremal values (in terms of the Problem statements A-D) of the objective (minimized) function $\varphi(y)$  over the domain $Q$ is defined by the \textit{coordinate} (\ref{eq:1_10}) and \textit{functional} (\ref{eq:1_9}) \textit{constraints} on the selection of feasible points (or vectors) $y=(y_1,\ldots ,y_N)$ .
+In this model, the constants $a_i,b_i, 1\leq i\leq N$,  determine the limits for the variation of the problem parameters (the components of the vector $y$), and are finite or (when the corresponding upper or/and lower limits are absent for some of them) they are assumed to be equal to $a_i=-\infty $  and/or $b_i=+\infty $.
 
@@
 Second, in numerical solving an optimization problem on a computer the exact calculation of the equality $h(y)=0$ is often impossible, therefore, this equality is assumed to be computed  with some error $\delta>0$ , i.e., the inequality $\left|h(y)\right|\leq \delta$ is considered instead of the equation $h(y)=0$.
-Thus, taking into account the above reasons, one can claim that (\ref{eq:8})--(\ref{eq:10}) is the statement of  \textit {the general  problem of the nonlinear programming}.
+Thus, taking into account the above reasons, one can claim that (\ref{eq:1_8})--(\ref{eq:1_10}) is the statement of  \textit {the general  problem of the nonlinear programming}.
 
-If $m=0$ , i.e., there are no functional constraints, let us assume $Q=D$. The problem (\ref{eq:8})--(\ref{eq:10}) in this case will be called \textit {the unconstrained optimization problem}.
+If $m=0$ , i.e., there are no functional constraints, it is considered that $Q=D$.
+The problem (\ref{eq:1_8})--(\ref{eq:1_10}) in this case will be called \textit {the unconstrained optimization problem}.
 
@@
 \begin{displaymath}
-%\label{eq:12}
+%\label{eq:1_12}
 \varphi(y_1,y_2)=-1.5y_1^2\exp(1-y_1^2-20.25(y_1-y_2)^2)-
@@
 \begin{displaymath}
-%\label{eq:12}
+%\label{eq:1_12}
 -(0.5(y_1-1)(y_2-1))^4\exp(2-(0.5(y_1-1))^4-(y_2-1)^4)
@@
 
-The set of the points satisfying the first constraint represents a circle with the boundary defined by the equation $g_1(y_1,y_2)=0$.
-The points satisfying the second constraint are located outside an ellipse defined by the equation $g_2(y_1,y_2)=0$.
+The set of the points satisfying the first constraint represents the circle with the boundary defined by the equation $g_1(y_1,y_2)=0$.
+The points satisfying the second constraint are located outside the ellipse defined by the equation $g_2(y_1,y_2)=0$.
 The points satisfying the third constraint are below the sinusoid $g_3(y_1,y_2)=0$.
@@
 In this case, the analytical investigation methods are unsuitable.
-Note that even if the function is defined analytically and if the derivative calculation is possible the initial problem (\ref{eq:6}) is reduced to the problem of finding all the derivative roots, the difficulty of which is comparable with the minimization problem.
+Note that even if the function is defined analytically and if the derivative calculation is possible the initial problem (\ref{eq:1_6}) is reduced to the problem of finding all the derivative roots, the difficulty of which is comparable with the minimization problem.
 
@@
 
-Let us give a formal definition of sequential optimization algorithm based on the methodology of the theory of operations research, or, more general, on the computational model for solving the problem (\ref{eq:6}).
+Let us give a formal definition of sequential optimization algorithm based on the methodology of the theory of operations research, or, more general, on the computational model for solving the problem (\ref{eq:1_6}).
 The generalization of this model onto the parallel computing case will be given in Subsect.
@@
 \begin{equation}
-\label{eq:11}
+\label{eq:1_11}
 s=\left\langle \{G_k\},\{E_k\},\{H_k\}\right\rangle
@@
 
--  $\{E_k\}$	  is a set of functionals defining the rules of constructing the approximate solution (the estimate of extremum), $k=1,2,\ldots$ ;
+-  $\{E_k\}$	  is a set of functionals defining the rules of constructing the approximate solution (the estimates of extremum), $k=1,2,\ldots$ ;
 
@@
 \begin{equation}
-\label{eq:12}
+\label{eq:1_12}
 y^1=G_1(\Phi\in Q)
@@
 is selected as the first trial coordinate and the trial counter $k=1$ .}
-\item[\textbf{Step 2}]{Assume the point of the $k$-th trial $y^k\in Q$  $(k\geq 1)$ has been selected.
+\item[\textbf{Step 2}]{Assume that the point of the $k$-th trial $y^k\in Q$  $(k\geq 1)$ has been selected.
 The value of the function $z^k=\varphi(y^k)$  is computed.
@@
 \begin{equation}
-\label{eq:13}
+\label{eq:1_13}
 \omega_k=\{(y^1,z^1),(y^2,z^2),\ldots ,(y^k,z^k)\}.
@@
 \begin{equation}
-\label{eq:14}
+\label{eq:1_14}
 \Phi(\omega_k)=\{\psi\in\Phi:\psi(y^i)=z^i,1\leq i \leq k\}.
@@
 \begin{equation}
-\label{eq:15}
+\label{eq:1_15}
 e^k=E_k(\Phi,\omega_k)
@@
 \begin{equation}
-\label{eq:16}
+\label{eq:1_16}
 y^{k+1}=G_{k+1}(\Phi,\omega_k)
@@
 \begin{equation}
-\label{eq:17}
+\label{eq:1_17}
 h^k=H_k(\Phi,\omega_k,y^{k+1})\in \{0,1\}
@@
 
-\example {Let us consider the simplest method of solving a one-dimensional problem (\ref{eq:6}) within an interval $[a,b]$, namely, the method of scanning  the function values in the nodes of a regular grid.
+\example {Let us consider the simplest method of solving a one-dimensional problem (\ref{eq:1_6}) within an interval $[a,b]$, namely, the method of scanning  the function values in the nodes of a regular grid.
 The method consists in the partitioning of the search interval $[a,b]$ into $q$ equal parts.
@@
 The computability of the function at any point of the search domain is sufficient for the applicability of the method.
-So, the class of the functions defined over the interval $[a,b]$ and being computable at every point of the interval could be considered as the a priori known class $\Phi$.
+So, the class of the functions defined over the interval $[a,b]$ and being computable at every point of the interval can be considered as the a priori known class $\Phi$.
 
@@
 \begin{displaymath}
-G_1(\Phi)=a, G_{k=1}(\Phi,\omega_k)=a+k\frac{b-a}{q}, k\geq 1,
+G_1(\Phi)=a, G_{k+1}(\Phi,\omega_k)=a+k\frac{b-a}{q}, k\geq 1,
 \end{displaymath}
@@
 \begin{equation}
-\label{eq:18}
+\label{eq:1_18}
 e^k=\varphi_k^*,
@@
 \begin{equation}
-\label{eq:19}
+\label{eq:1_19}
 \varphi_k^*=\min_{1\leq i\leq k} \varphi(y^i),
@@
 \begin{displaymath}
-%\label{eq:19}
+%\label{eq:1_19}
 y_k^*=\arg \min_{1\leq i\leq k} \varphi(y^i).
@@
 \label {subsec:1.2.2}
-The computational scheme (\ref{eq:12})--(\ref{eq:17}) implies  the sequential execution of all the algorithm steps, that can be provided by the single processor computers.
+The computational scheme (\ref{eq:1_12})--(\ref{eq:1_17}) implies  the sequential execution of all the algorithm steps, that can be provided by the single processor computers.
 But if a multiprocessor computer is available, a parallelization of the computational scheme and, therefore, an acceleration of the optimization process can be realized.
 The parallelization of optimization methods could be done in several ways.
-It is possible to parallelize, in the first place, the calculation of the objective function, secondly, the implementation of the rules $G_k,E_k,H_k$ , and, finally, the scheme (\ref{eq:12})--(\ref{eq:17}) providing simultaneous execution of several trials.
+It is possible to parallelize, in the first place, the calculation of the objective function, secondly, the implementation of the rules $G_k,E_k,H_k$ , and, finally, the scheme (\ref{eq:1_12})--(\ref{eq:1_17}) providing simultaneous execution of several trials.
 However, the parallelization of the objective function describing the optimized object is specific for different objective function models, and as a matter of fact, cannot be related to the general optimization scheme.
@@
 Besides, often these functionals are rather simple, and their parallelization is just of no use.
-Therefore, Therefore, the subject of the consideration in the present subsection  is the design of some \textit {fundamental principles} for a parallel execution of trials.
+Therefore, the subject of the consideration in the present subsection  is the consideration of some \textit {fundamental principles} for a parallel execution of trials.
 Let us use the term \textit {iteration} for the simultaneous (parallel) execution of several trials (each trial is executed by a separate processor).
@@
 \begin{equation}
-\label{eq:20}
-y^1=G_1(\Phi\in Q),y^i=G_i(\Phi,u_{i-1}(1))\in Q,\ 2\leq i\leq p(1)
+\label{eq:1_20}
+y^1=G_1(\Phi)\in Q,y^i=G_i(\Phi,u_{i-1}(1))\in Q,\ 2\leq i\leq p(1)
 \end{equation}
 and assign the number of trials $k(1)=p(1)$.}
-\item[\textbf{Step 2}]{Assume that the points of the $n$th iteration $u(n),\ n\geq 1,$ , have been obtained.
-The computations of the objective function values $z^j=\varphi(y^j),\ k(n-1)+1\leq j\leq k(n)$, are executed (each value is computed by a separate processor).
+\item[\textbf{Step 2}]{Assume that the points of the $n$-th iteration $u(n),\ n\geq 1,$ , have been obtained.
+The computations of the objective function values $z^j=\varphi(y^j)$,   $\ k(n-1)+1\leq j\leq k(n)$, are executed (each value is computed by a separate processor).
 After computing all the values there is the following search information on the function $\varphi(y)$ (\textit{a posteriori information}):
 \begin{equation}
-\label{eq:21}
+\label{eq:1_21}
 \omega_k=\omega_{k(n)}=\{(y^1,z^1),(y^2,z^2),\ldots ,(y^{k(n)},z^{k(n)})\}.
 \end{equation}
-As before, the new information (\ref{eq:21}) allows one to reduce the class, which the function $\varphi(y)$ belongs to, up to the set   
+As before, the new information (\ref{eq:1_21}) allows one to reduce the class, which the function $\varphi(y)$ belongs to, up to the set   
 \begin{equation}
-\label{eq:22}
+\label{eq:1_22}
 \Phi(\omega_{k(n)})=\{\psi\in\Phi:\psi(y^i)=z^i,1\leq i \leq k(n)\}.
@@
 \begin{equation}
-\label{eq:23}
+\label{eq:1_23}
 e^{k(n)}=E_{k(n)}(\Phi,\omega_{k(n)}).
@@
 }
-\item[\textbf{Step 4}]{$p(n+1)$   coordinates of the trials of the $n+1$-th iteration are computed:
+\item[\textbf{Step 4}]{Coordinates of the trials at the ($n+1$)-th iteration are computed:
 \begin{equation}
-\label{eq:24}
+\label{eq:1_24}
 \begin{gathered}
 y^{k(n)+1}=G_{k(n)+1}(\Phi,\omega_{k(n)}) \\
-y^i=G_i(\Phi,\omega_{k(n)},u_{i-1}(k(n+1))
+y^{k(n)+i}=G_{k(n)+i}(\Phi,\omega_{k(n)},u_{i-1}),\  2\leq i\leq p(n+1),
 \end{gathered}
@@
 \begin{equation}
-\label{eq:25}
-h^{k(n)}=H_{k(n)}(\Phi,\omega_{k(n)},u_{n+1})\in \{0,1\}
+\label{eq:1_25}
+h^{k(n)}=H_{k(n)}(\Phi,\omega_{k(n)},u(n+1))\in \{0,1\}
 \end{equation}
 taking one of two possible values: 0 or 1 is determined.
-If $h^{k(n)}=1$ the iteration counter $k$ is incremented ($n=n+1$), and the return to the Step 2 of the scheme is carried out.
+If $h^{k(n)}=1$ the iteration counter $n$ is incremented ($n=n+1$), and the return to the Step 2 of the scheme is carried out.
 If $h^{k(n)}=0$ the  iterations are completed, and the estimate $e^{k(n)}$  is taken as the (approximate) problem solution.}
@@
 \end{displaymath}
-is determined and $p(n+1)$  trial coordinates of the current iteration are calculated according to (\ref{eq:24}).}
+is determined and $p(n+1)$  trial coordinates of the current iteration are calculated according to (\ref{eq:1_24}).}
 \end{description}
@@
 Let us assume for simplicity that the number of processors $p$ is the same for all iterations, that the number of nodes $q>p$  and is divisible by $p$.
-Then, the decision rules $G(\bullet )$  in the expressions (\ref{eq:20}) and (\ref{eq:24})  obviously take the form
+Then, the decision rules $G(\bullet )$  in the expressions (\ref{eq:1_20}) and (\ref{eq:1_24})  obviously take the form
 \begin{displaymath}
-G(\bullet )=a+(i-1)\frac{b-a}{q},\ 1\leq i\leq k(n)+1,
+G(\bullet )=a+(j-1)\frac{b-a}{q},\ k(n)+1\leq j\leq k(n)+p(n+1),
 \end{displaymath}
@@
 Therefore, it is convenient to deal with the number of trials $k$.
-In this connection, let us introduce a set of completed trials  $w(k)=\{w_1,w_2,\ldots,w_k\}$  and a set  of coordinates $v(k)=\{v_1,v_2,\ldots,v_{p(k)}\}$ , at which new trials are executed after completing the preceding $k$ trials.
+Thereby, let us introduce a set of completed trials  $w(k)=\{w_1,w_2,\ldots,w_k\}$  and a set  of coordinates $v(k)=\{v_1,v_2,\ldots,v_{p(k)}\}$ , at which new trials are executed after completing the preceding $k$ trials.
 Here $p(k)$ is the number of processors executing the current trials, and $w_j=y^{i_j},\ 1\leq j\leq k,\ v_s=y^{i_s},\ 1\leq s\leq p(k), 1\leq i_j,i_s \leq k+p(k)$ .
@@
 \begin{equation}
-\label{eq:26}
+\label{eq:1_26}
 \begin{gathered}
@@
 \begin{equation}
-\label{eq:27}
+\label{eq:1_27}
 \omega_k=\{(y^{i_1},z^{i_1}),(y^{i_2},z^{i_2}),\ldots ,(y^{i_k},z^{i_k}\}.
@@
 \begin{equation}
-\label{eq:28}
+\label{eq:1_28}
 e^k=E_k(\Phi,\omega_k).
@@
 \begin{equation}
-\label{eq:29}
+\label{eq:1_29}
 \begin{gathered}
 y^{k+1}=G_{k+1}(\Phi,\omega_{k}) \\
-y^i=G_i(\Phi,\omega_k,y^{k+1},\ldots,y^{k+i-1}),\;k+2\leq i\leq k+\pi,
+y^{k+i}=G_i(\Phi,\omega_k,y^{k+1},\ldots,y^{k+i-1}),\;2\leq i\leq \pi,
 \end{gathered}
@@
 \begin{equation}
-\label{eq:30}
+\label{eq:1_30}
 h^k=H_k(\Phi,\omega_k,y^{k+1},\ldots,y^{k+\pi})\in \{0,1\}
@@
 
-Some examples of both the synchronous and asynchronous optimization algorithms are presented in the next chapters.
+Some examples of both the synchronous and asynchronous optimization algorithms are presented can be found in this monograph (see \cite{1_BarGerStrChap4, 1_BarGerStrChap6, 1_GriIsrChap5, 1_GriSergChap3}).
 \subsection {Convergence and estimates of extremum}
@@
 At the same time, since a computer can execute a finite number of trials only, it is desirable to obtain the exact solution after constructing a finite sequence $\{y^k\}$.
-Therefore,  an asymptotically exact assessment is interesting often, and an infinite trial sequence is considered (in the introduced  optimization method models in (\ref{eq:17}) , (\ref{eq:25}), (\ref{eq:30}) $H_k=1$  for any $k\geq1$, i.e., in fact, the termination criterion is absent).
-This sequence is required to converge to the exact problem solution.
-Since in the problem statements $B-D$ the desired solution may include several minimum points, we will understand the convergence of a method in the scope of the following definition.
+Unfortunately, such a favorable situation takes place in rare and simple cases only, for example, in the linear programming problems.
+Therefore,  an asymptotically exact assessment is interesting often, and an infinite trial sequence is considered (in this case for the introduced  optimization method models in (\ref{eq:1_17}) , (\ref{eq:1_25}), (\ref{eq:1_30}) $H_k=1$  for any $k\geq1$, i.e., in fact, the termination criterion is absent).
+In other words, this sequence is required to converge to the exact problem solution.
+Since in the problem statements $B-D$ the desired solution may include several minimum points, we will understand the convergence of a method in the sense of the following definition.
 \begin{definition} 
 \label{def:1_4}
-The trial sequence $\{y^k\}$  converges to the solution of the problem (\ref{eq:6}) defined by the corresponding problem statement, if:
+The trial sequence $\{y^k\}$  converges to the solution of the problem (\ref{eq:1_6}) defined by the corresponding problem statement, if:
 \begin{enumerate}
@@
 Assume that the values of the function $\varphi(y)$ are calculated in a finite number of points   What is it possible to say about the coordinate of the global minimum after that?
-Whatever points $y^1,y^2,\ldots,y^k$  and values $z^1,z^2,\ldots,z^k$ , were obtained after $k$  trials, for any point $y^*\in Q (y^*\notin \{y^1,y^2,\ldots,y^k\})$   a continuous function passing through the points $(y^i,z^i),1\leq i\leq k$,  i.e., belonging to the class $\Phi(\omega_k$ from (\ref{eq:14}) and providing the global minimum at the point $y^*$  with any predefined value
+Whatever points $y^1,y^2,\ldots,y^k$  and values $z^1,z^2,\ldots,z^k$ were obtained after $k$  trials, for any point $y^*\in Q \ \ (y^*\notin \{y^1,y^2,\ldots,y^k\})$   a continuous function passing through the points $(y^i,z^i),1\leq i\leq k$,  i.e., belonging to the class $\Phi(\omega_k)$ from (\ref{eq:1_14}) and providing the global minimum at the point $y^*$  with any predefined value
 \begin{displaymath}
@@
 All the above means that no conclusions on the position of the global minimum point can be drawn after any  finite number of trials.
-Similarly, the only conclusion on the value $\varphi^*$  of the global minimum could be as follows:
+Similarly, the only conclusion on the value $\varphi^*$  of the global minimum can be as follows:
 \begin{displaymath}
@@
 \end{displaymath}
-where $\varphi_k^*$  is determined by (\ref{eq:19}). Therefore, it is impossible to estimate the value 
+where $\varphi_k^*$  is determined by (\ref{eq:1_19}). Therefore, it is impossible to estimate the value 
 \begin{displaymath}
@@
 This interesting problem falls out of the scope of the present paper.
-However, the availability of the extremum estimates is important for the investigation of the parallelization efficiency of the optimization methods considered in the next chapters.
+However, the availability of the extremum estimates is important for the investigation of the parallelization efficiency of the optimization methods.
 
@@
 In multiextremal problems the dimension affects the difficulty of solving such problems significantly.
-For example, for the class of the functions satisfying Lipschitz condition so called «the curse of dimensionality» consisting in the exponential growth of the computational costs with increasing dimensionality takes place.
+For example, for the class of the functions satisfying Lipschitz condition so called  \textit{curse of dimensionality} consisting in the exponential growth of the computational costs with increasing dimensionality takes place.
 Namely, if $p$ evaluations of the objective function are required to achieve the solution accuracy $\epsilon$ in the one-dimensional case, in the $N$-dimensional problem $\alpha p^N$  trials are required to achieve the same accuracy.
@@
 
-Another approach to the design of the numerical methods for the analysis of the multidimensional multiextremal problems exploits the idea of \textit{the complexity reduction} where the solution of the initial problem is replaced by solving one or several simpler problems.
+Another approach to the design of the numerical methods for the analysis of the multidimensional multiextremal problems exploits the idea of \textit{the complexity reduction} where the solution of the initial problem is replaced with solving one or several simpler problems.
 
@@
 In the context of the theoretical approach this construct seems to be a perfect one.
-However, in practice everything is not so rosy.
+However, in practice some difficulties arise.
 First of all, the following question arises: how to find all local minima?
-This problem could be solved, if for each local minimum its attraction zone, i.e., such neighborhood of the minimum point, in which the function is unimodal, is known.
+This problem can be solved, if for each local minimum its attraction zone, i.e., such neighborhood of the minimum point, in which the function is unimodal, is known.
 Then, having positioned the initial search point into this neighborhood, one can find the desired local minimum by some local method.
@@
 
-The convergence to the global extremum in this method is provided by the fact that if to increase the number of the starting points at least one of them should fall into the global extremum attraction zone since these grids (either regular or random) are built to provide a coverage of the whole search domain with increasing density.
+The convergence to the global extremum in this method is provided by the fact that if to increase the number of the starting points at least one of them will fall into the global extremum attraction zone since these grids (either regular or random) are built to provide a coverage of the whole search domain with increasing density.
 However, this scheme has an essential drawback.
 As a matter of fact, several starting points can fall into the attraction zone of a local minimum, i.e., this particular local minimum will be found several times.
-Different schemes have been proposed to overcome this disadvantage ( see \cite{BetroSchoen, BoendeRinnooy, 1_Zielinski}).
+Different schemes have been proposed to overcome this disadvantage ( see \cite{1_BetroSchoen, 1_BoendeRinnooy, 1_Zielinski}).
 
@@
 \begin{equation}
-\label{eq:31}
+\label{eq:1_31}
 \lim_{\tau \to \infty}\min_{1\leq k \leq \tau}{\varphi(y^k)}=\varphi^*.
@@
 Avoiding this drawback is possible only if enough and rather rich a priori known information on the problem is available (like the data on the attraction zones).
-Such information is rather rarely available in practice and sometimes could be obtained for simple multiextremal problems with a few number of extrema only.
+Such information is rather rarely available in practice and sometimes can be obtained for simple multiextremal problems with a few number of extrema only.
 For such problems this approach is quite efficient.
 However, for the essentially multiextremal problems the application of reduction to the local minimum problems is not successful, as a rule.
-At the same time, the methods of this class are parallelized rather easily, but the everywhere dense convergence (\ref{eq:31}) can results in a large number of redundant trials.
+At the same time, the methods of this class are parallelized rather easily, but the everywhere dense convergence (\ref{eq:1_31}) can results in a large number of redundant trials.
 
@@
 \item{the nested optimization scheme  of optimization \cite{1_CarrHowe, 1_Evtushenko, 1_GerGriIsr, 1_GerGriGer, 1_GriIsrAIP, 1_GriIsrSerg, 1_GriIsrCEUR, 1_Piyavskij, 1_SergGriJCAA, 1_ShiOlaf, 1_StrMonRus, 1_StrSergMon2000, 1_vanDam},}
-\item{the reduction of dimensionality based on the curves filling the space (Peano-type curves) \cite{1_Butz, 1_Goertzel, 1_HimOliPet, 1_LeraSergCNSNS, 1_LeraSergANM, 1_SergStrLeraMonogr, 1_StrMonRus,1_StrSergMon2000}.}
+\item{the reduction of dimensionality based on the curves filling 
+space (Peano-type curves) \cite{1_Butz, 1_Goertzel, 1_HimOliPet, 1_LeraSergCNSNS, 1_LeraSergANM, 1_SergStrLeraMonogr, 1_StrMonRus,1_StrSergMon2000}.}
 \end{enumerate}
-The \textit{nested optimization scheme} of dimensionality reduction replaces solving the problem (\ref{eq:8})--(\ref{eq:10}) with solving a family of recursively nested one-dimensional subproblems, in which every evaluation of the objective function within a one-dimensional subproblem consists in solving a new one-dimensional subproblem of the next level of the recursion except the last one, where the value of the initial multidimensional objective function is calculated.
+The \textit{nested optimization scheme} of dimensionality reduction replaces solving the problem (\ref{eq:1_8})--(\ref{eq:1_10}) with solving a family of recursively nested one-dimensional subproblems, in which every evaluation of the objective function within a one-dimensional subproblem consists in solving a new one-dimensional subproblem of the next level of the recursion except the last one, where the value of the initial multidimensional objective function is calculated.
 
 The second scheme is the dimensionality reduction based on the Peano-type curves.
-It is based on the well known fundamental fact, according to which the $N$-dimensional hyperparallelepiped (\ref{eq:10}) and the interval $[0,1]$ of the real axis are the equinumerous sets, and the interval $[0,1]$ can be mapped onto the parallelepiped (\ref{eq:10}) , unambiguously and continuously \cite{1_SergStrLeraMonogr, 1_StrSergMon2000}.
+It is based on the well known fundamental fact, according to which the $N$-dimensional hyperparallelepiped (\ref{eq:1_10}) and the interval $[0,1]$ of the real axis are the equinumerous sets, and the interval $[0,1]$ can be mapped onto the parallelepiped (\ref{eq:1_10}) , unambiguously and continuously \cite{1_SergStrLeraMonogr, 1_StrSergMon2000}.
 The representations of this kind are usually called \textit{Peano-type curves} or \textit{evolvents}.
 
-Let $y(x),[\in [0,\:1]$  be a Peano-type curve, and the function  $\varphi(y)$ from (\ref{eq:8}) be continuous. Then because of the continuity of  $\varphi(y)$, $y(x)$ and owing to the relation
+Let $y(x),[\in [0,\:1]$  be a Peano-type curve, and the function  $\varphi(y)$ from (\ref{eq:1_8}) be continuous. Then because of the continuity of  $\varphi(y)$, $y(x)$ and owing to the relation
 \begin{displaymath}
@@
 
-The consideration and justification of the algorithms based on the nested optimization scheme and on the dimensionality reduction schemes using Peano-type curves is the major subject of the next chapters.
+The detailed consideration and justification of the algorithms based on the nested optimization scheme and on the dimensionality reduction schemes using Peano-type curves can be found, for example, in this monograph (see \cite{1_BarGerStrChap6, 1_GriIsrChap5}).
 % 
 \begin{thebibliography}{99.}
-\bibitem{BetroSchoen} Betro, B., Schoen, F.: Sequential stopping rules for the multistart algorithm in global optimisation. Math. Program. \textbf{38}(3), 271--286 (1987)
-\bibitem{BoendeRinnooy}	Boender, C.G.E, Rinnooy Kan, A.H.G.: Bayesian stopping rules for multistart global optimization methods. Math. Program. \textbf{37}(1), 59--80 (1987)
+\bibitem{1_BarGerStrChap4} Barkalov, K.A., Gergel, V.P., Strongin, R.G.: Global search algorithms for one-dimensional multiextremal optimization problems with nonlinear constraints. This volume.
+\bibitem{1_BarGerStrChap6} Barkalov, K.A., Gergel, V.P., Strongin, R.G.: Parallel computations for the multidimensional multiextremal optimization problems with the dimensionality reduction using Peano curves.
+This volume.
+\bibitem{1_BetroSchoen} Betro, B., Schoen, F.: Sequential stopping rules for the multistart algorithm in global optimisation. Math. Program. \textbf{38}(3), 271--286 (1987)
+\bibitem{1_BoendeRinnooy}	Boender, C.G.E, Rinnooy Kan, A.H.G.: Bayesian stopping rules for multistart global optimization methods. Math. Program. \textbf{37}(1), 59--80 (1987)
 \bibitem{1_Butz}	Butz, A.R. Space-Filling Curves and Mathematical Programming.: Inform. Control \textbf{12}, 314--330 (1968)
@@
 \bibitem{1_Goertzel}	Goertzel, B.: Global Optimization with Space-Filling Curves. Appl. Math. Lett. 12, 133--135 (1999)
+\bibitem{1_GriIsrChap5}Grishagin, V.A., Israfilov, R.A.: Sequential and parallel algorithms for multidimensional multiextremal optimization based on the nested scheme of dimensionality
+reduction. This volume.
 \bibitem{1_GriIsrAIP}	Grishagin, V.A., Israfilov, R.A.: Global search acceleration in the nested optimization scheme. AIP Conf. Proc. \textbf{1738}, 400010 (2016)
@@
 (2017) doi:10.1016/j.amc.2017.06.036
-\bibitem{1_GriIsrCEUR}Grishagin, V.A., Israfilov, R.A.: Multidimensional Constrained Global Optimization in Domains with Computable Boundaries. CEUR Workshop Proceedings 1513, 75--84 (2015)
+\bibitem{1_GriIsrCEUR} Grishagin, V.A., Israfilov, R.A.: Multidimensional Constrained Global Optimization in Domains with Computable Boundaries. CEUR Workshop Proceedings 1513, 75--84 (2015)
+\bibitem{1_GriSergChap3} Grishagin, V.A., Sergeyev, Y.D.: Parallel algorithms for one-dimensional multiextremal optimization. This volume.
 \bibitem{1_HimOliPet}	Hime, A.E., Oliveira Jr., H.A., Petraglia, A.: Global Optimization Using Space-Filling Curves and Measure-Preserving Transformations.
@@
 \end{thebibliography}
-%\input{referenc}
+
 %\end{document}

### 2017-10-07 2017/ParOptBook/Chapter6.tex
Chapter -> paper

@@
 
-\abstract*{The chapter considers dimensionality reduction scheme based on Peano-type space-filling cures (evolvents).
+\abstract*{The paper considers dimensionality reduction scheme based on Peano-type space-filling cures (evolvents).
 The initial multidimensional optimization problem is substituted by an equivalent one-dimensional one based on the application of the mapping of a multidimensional search domain onto a segment of the real axis.
@@
 
-\abstract{The chapter considers dimensionality reduction scheme based on Peano-type space-filling cures (evolvents).
+\abstract{The paper considers dimensionality reduction scheme based on Peano-type space-filling cures (evolvents).
 The initial multidimensional optimization problem is substituted by an equivalent one-dimensional one based on the application of the mapping of a multidimensional search domain onto a segment of the real axis.

### 2017-10-07 2017/ParOptBook/Chapter8.tex
Chapter -> paper

@@
 
-\abstract*{In this chapter, the examples of solving the applied problems in the fields of optimal design (the finding of the optimum profile of a vehicle wheel), chemical research (finding the unknown kinetic parameters in the mathematical models of the complex chemical reactions), economics (the identification of the parameters in the regional economic model) are given.}
+\abstract*{In this paper, the examples of solving the applied problems in the fields of optimal design (the finding of the optimum profile of a vehicle wheel), chemical research (finding the unknown kinetic parameters in the mathematical models of the complex chemical reactions), economics (the identification of the parameters in the regional economic model) are given.}
 
-\abstract{In this chapter, the examples of solving the applied problems in the fields of optimal design (the finding of the optimum profile of a vehicle wheel), chemical research (finding the unknown kinetic parameters in the mathematical models of the complex chemical reactions), economics (the identification of the parameters in the regional economic model) are given.}
-
-%\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\abstract{In this paper, the examples of solving the applied problems in the fields of optimal design (the finding of the optimum profile of a vehicle wheel), chemical research (finding the unknown kinetic parameters in the mathematical models of the complex chemical reactions), economics (the identification of the parameters in the regional economic model) are given.}
 
@@
 The typical computation costs were several days of continuous computations.
-Within the framework of the joint research project by IPC RAS and Lobachevsky State University of Nizhni Novgorod it was proposed to apply the efficient parallel global optimization methods described in Chapter 5.
-%Ссылка на главу!!!!
+Within the framework of the joint research project by IPC RAS and Lobachevsky State University of Nizhni Novgorod it was proposed to apply the efficient parallel global optimization methods developed in UNN.
 
@@
 
-\section*{A brief conclusion}
-
-In this chapter, some applied global optimization problems have been considered.
-All of these problems related with a search for optimal parameters for nonlinear relationships, which describe phenomenon or process under consideration.
-The first example is the finding of the optimum profile of a vehicle wheel.
-Solution to this problem allows to increase the maximum possible speed of the vehicle.
-The second one is related with chemical research (finding the unknown kinetic parameters in the mathematical models of the complex chemical reactions).
-The last example belongs to the field of economics.
-The problem of identification of the parameters for the regional economic model has been solved.
-Optimal parameters found with the global search algorithm allow to carry out scenario forecast calculations for the regional economy.
-
 \begin{thebibliography}{99.}

### 2017-10-09 2017/ParOptBook/Chapter7.tex
Chapters 7 and 8 are merged. Chapter 8 is not necessary now.

@@
 %
-% sample root file for your "contribution" to a contributed volume
+%% sample root file for your "contribution" to a contributed volume
 %
-% Use this file as a template for your own input.
+%% Use this file as a template for your own input.
 %
 %%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %\documentclass[graybox]{svmult}
-%
-%% choose options for [] as required from the list
-%% in the Reference Guide
-%
+%%
+%%% choose options for [] as required from the list
+%%% in the Reference Guide
+%%
 %\usepackage{mathptmx}       % selects Times Roman as basic font
@@
 %\usepackage{courier}        % selects Courier as typewriter font
-%\usepackage{type1cm}        % activate if the above 3 fonts are
-                             % not available on your system
-%
+%%\usepackage{type1cm}        % activate if the above 3 fonts are
+                             %% not available on your system
+%%
 %\usepackage{makeidx}         % allows index generation
 %\usepackage{graphicx}        % standard LaTeX graphics tool
-%                             % when including figure files
+%%                             % when including figure files
 %\usepackage{multicol}        % used for the two-column index
 %\usepackage[bottom]{footmisc}% places footnotes at page bottom
+%
+%
+%\usepackage[colorlinks=true]{hyperref}
+%\hypersetup{urlcolor=blue, citecolor=blue}
+%
 %
@@
 %\makeindex             % used for the subject index
-%                       % please use the style svind.ist with
-%                       % your makeindex program
+                       %% please use the style svind.ist with
+                       %% your makeindex program
 %
@@
 
-\title{Software Systems for Global Optimization }
-% Use \titlerunning{Short Title} for an abbreviated version of
-% your contribution title if the original one is too long
-\author{Name of First Author and Name of Second Author}
-% Use \authorrunning{Short Title} for an abbreviated version of
-% your contribution title if the original one is too long
-\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
-\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
+\title{Software systems for global optimization and the applied optimization problems}
+\titlerunning{Software systems for global optimization }
+
+\author{Victor Gergel, Konstantin Barkalov, Alexander Sysoyev }
+\authorrunning{V.Gergel, K. Barkalov, A. Sysoyev}
+
+\institute{Victor Gergel, Konstantin Barkalov, Alexander Sysoyev  \at Lobachevsky State University of Nizhni Novgorod,  Nizhni Novgorod, Russia \email{konstantin.barkalov@itmm.unn.ru}}
 %
@@
 
-\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\abstract{This paper contains a description of the program systems for solving global optimization problems which have been developed by the authors.
+The first two systems -- GlobaLab and ParaLab -- are intended for educational and research goals and allow one to study both the known and new global search algorithms.
+These systems enable to realize the statement of an optimization problem, to choose a method of its solving, to execute a computational experiment and to analyze the results obtained.
+Such the capacities allow assimilating an experience in solving various global optimization problems, evaluating their efficiency and getting necessary knowledge and skills.
+After the study of the theory and practical training in optimization on the base of the abovementioned systems the reader can get acquainted with the GlobalExpert system that provides the possibility of solving complicated optimization problems arising in different areas of applications.
+The examples of solving the applied problems in the fields of optimal design, chemical research and economics are given.}
 
-\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\section{The software system GlobaLab for investigation and study of global optimization methods}
 
-\section{Section Heading}
-Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.
+Globalizer Laboratory (\textit{GlobaLab}) is an integrated software system for carrying out the experiments with the global search methods for study and investigation of the basic concepts, approaches, and the algorithms in the field of \textbf{global optimization}.
+
+The multiextremal optimization problems considered within the framework of the theory of global search are the subject of extensive studies and are applied widely in computer aided design, in solving the identification problems, etc.
+GlobaLab belongs to the class of systems for carrying out the computational experiments, the development of which is an important area of the novel research techniques in various fields of science and technologies.
+
+The availability of all necessary tools for carrying out the computational experiments is a distinguishing feature of the system.
+Using GlobaLab, the user of the system (\textit{the optimizer}) can formulate the statement of an optimization problem, select the optimizing algorithm from the list of the global search methods available in the system, set up the graphic indicators for monitoring the global search process, carry out various computational experiments, obtain and analyze the experiment results.
+These completeness and orientation on the novel technologies of education distinguish GlobaLab from other systems (such as, for example, SYMOP \cite{7_Gergel1993} and LGO \cite{7_Pinter1996}).
+
+The possible fields of application of the system are:
+\begin{itemize}
+\item \textit{education}: applying by the students and teachers for their investigations and study of the global search methods within the framework of the laboratory training for various disciplines of the optimization oriented curriculum (optimization methods, system analysis, operations research, etc.).
+GlobaLab can be used for demonstrations of a number of important concepts of mathematics and computer science (sequences, limit points, convergence of a numerical method, etc.).
+GlobaLab can also be used as an example of a complex approach to the development of the computer simulation systems;
+
+\item \textit{research}: applying by the researchers in the fields of the multiextremal optimization theory for the investigations of the efficiency of the global search algorithms under development.
+The system can be used as a tool for preparing and carrying out the scientific demonstrations of numerical solving the optimization problems by various methods;
+
+\item \textit{applications}: applying by the experts and researchers as a training component of the complex professional optimization systems.
+\end{itemize}
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=1.0\linewidth]{figures/7_1.png}
+\caption{GlobaLab main window}
+\label{7_fig_1}     
+\end{figure}
+
+The users beginning to study the global search problems can find GlobaLab to be useful to get acquainted with the multiextremal optimization methods.
+The experienced optimizers can use the system to estimate the efficiencies of newly developed global search algorithms.
+A screenshot of the GlobaLab main window is presented in Fig.~\ref{7_fig_1}.
+
+\textbf{Capacities of the system.} GlobaLab is an integrated program environment for carrying out the experiments with the global search methods. The user of the system is allowed:
+
+\begin{itemize}
+\item \textit{to formulate the statement of the optimization problem}.
+The minimized function can be selected from a list of the standard functions, generated by a random generator, defined by a formula, or formed by a graphic editor;
+
+\item \textit{to select a global search algorithm} from the list of the optimization methods implemented in the system.
+GlobaLab supports 10 various multiextremal algorithms, the most known in the theory and practice of the global search.
+The system provides a possibility to extend the set of methods by the build-in tools without using the algorithmic programming languages;
+
+\item \textit{to set up the graphic indicators} for monitoring the global search process; these indicators allow visualizing the graph of the minimized function or its piecewise linear approximation, the distribution, density, and sequence of the iteration points and of the function values at these points.
+Monitoring the indicators in the global search process provides better understanding of the global optimization theory and develops the intuition necessary for the practical application and further development of the multiextremal optimization methods;
+
+\item \textit{to carry out various computational experiments} sequentially or in parallel.
+The latter case allows visual comparing the search dynamics by various methods and is realized in the time sharing mode.
+Carrying out the series of computational experiments requiring long computation times can be performed automatically with the option of storing the search results for further analysis of the data obtained.
+Executing the numerical experiments can be performed in the manual mode as well, when the optimizer indicates the iteration points manually (this mode allows the user to verify various hypotheses, which can serve as the base for the development of novel global search methods);
+
+\item \textit{to accumulate and analyze the results of the computational experiments}.
+GlobaLab allows estimating the operation characteristics of the methods and provides the logging of the experiments in order to record the results of the optimization.
+The accumulated data can be presented in various generalized forms (tables, graphs, charts) convenient for further analysis.
+The results of the computations can be stored in a system archive, printed out, or copied to the Windows as text, graphics, or table and transferred to Word, Excel, or another Windows application programs for further processing and analysis (including for the use in preparing the scientific publications, reports, etc.).
+\end{itemize}
+
+\section{The software system ParaLab for studying and investigation of the parallel computation methods}
+
+Parallel Laboratory (\textit{ParaLab}) is a software system that provides the possibility to carry out the computational experiments with the purpose of studying and investigation of the parallel algorithms for solving complex computational problems.
+The system can be used in the laboratory training in various training courses in the field of parallel programming, within which the following possibilities are provided:
+
+\begin{itemize}
+\item \textit{the modeling of the parallel multiprocessor computation systems} with different topologies of the networks for the data transfer;
+
+\item \textit{the visualization} of the computational processes and of the data transfer operations at the parallel solving of different computational problems;
+
+\item \textit{the estimation of the efficiency} of the parallel computation methods studied.
+\end{itemize}
+
+The conduction of such training can be organized on the «ordinary» single processor computers running under the operation systems of MS Windows family (the multiprocess parallel computations simulation mode).
+Besides the simulation mode, ParaLab can provide the remote access to a multiprocessor computational system for carrying out the numerical experiments in the «true» parallel computations mode in order to compare the results of simulation with the ones of real parallel computations.
+
+In general, ParaLab system is an integrated computational environment for studying and investigation of the parallel algorithms for solving complex computational problems.
+A wide choice of available tools for the visualization of the computational experiment execution process and for the analysis of the obtained results allows studying the efficiency of application of various algorithms using various computational systems, to make the conclusions on the scalability of the algorithms, and to determine possible speedup for the parallel computation processes.
+
+The educational and research capabilities provided by ParaLab are directed onto the active learning of the theoretical basics and methods, onto encouraging the students to develop their own ideas on the models and methods of the parallel computations by visualization, comparison, and analysis of large set of various graphic indicators and viewers, activated during the experiments.
+
+The main application area of the system is the application in education by the students and teachers for the investigation and studying the parallel algorithms for solving the complex computational problems within the framework of the laboratory practicum in various training courses in the field of parallel programming.
+ParaLab system can be used also in carrying out research for the estimation of the parallel computations.
+The users, beginning to study the problems of parallel computing will find ParaLab system to be useful in learning the methods of parallel programming.
+The experienced programmers can use the system to estimate the efficiency of newly developed parallel algorithms.
+A screenshot of the main window of ParaLab system, which a test optimization problem was run in, is presented in Fig.~\ref{7_fig_2}).
+
+\textbf{Capacities of the system.} ParaLab is a software system, which allows conducting the real parallel computations on a multiprocessor computational system as well as the simulations of such experiments on a single serial computer with the visualization of the process of the parallel solving a complex computational problem.
+
+In carrying out the simulation experiments, ParaLab provides the following possibilities for the users:
+
+\begin{itemize}
+\item \textit{to define the topology} of the parallel computational system for carrying out the experiments, to define the number of processors in this topology, to set the performance of the processors, to select the parameters of the communication environment and the communication method;
+
+\item \textit{to formulate the statement of the computational problems}, for which in the system there are the built-in parallel algorithms to solve it, to set the parameters of the problem;
+
+\item \textit{to select a parallel method} for solving the specified problem;
+
+\item \textit{to set the visualization parameters} to choose the desired rate of the demonstration, the displaying method for the data transferred between the processors, and the level of the visualization details in the parallel computations;
+
+\item \textit{to carry out the computational experiments} on the parallel solving the specified problem; several different \textit{computational experiments} can be formed in ParaLab with different types of multiprocessor systems, problems, or parallel methods, which the experiments can be run in parallel (in the time sharing mode); the parallel execution of several experiments allows to compare the dynamics of the problem solving by various methods using various topologies with various parameters of the studied problem.
+If a series of experiments is executed requiring long-time computations, the system provides a possibility to execute the experiments in an automatic mode with saving the numerical results for further analysis of the obtained data,
+
+\item \textit{to accumulate and analyze the results of the experiments}; the system provides the tools for visualizing the saved data featuring the parallel computations (computation time, speedup, efficiency) as the functions of the parameters of the problem and of the computational system.
+\end{itemize}
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=1.0\linewidth]{figures/7_2.png}
+\caption{ParaLab main system (global optimization problem is selected)}
+\label{7_fig_2}     
+\end{figure}
+
+The possibility to choose the method of carrying out the experiments is one of the most important features of ParaLab.
+The experiments could be performed \textit{in the simulation mode}, i.e., performed on a single processor without using any parallel specialized software like the message transfer libraries.
+Besides, ParaLab provides the possibility to carry out \textit{the real computational experiments on parallel computers}.
+
+In visualizing the dependencies of the time characteristics on the parameters of the problem and of the computational system for the experiments carried out in the simulation mode the theoretical estimates are used according to the available models of the parallel computations.
+For the real computational experiments on the multiprocessor computer systems, the dependencies are visualized according to the results of the real numerical experiments.
+
+It is worth noting that ParaLab provides storing the results of the executed experiments in a special memory buffer.
+The stored results allow performing the analysis of the obtained data.
+Any experiment executed earlier can be restored to be recomputed again or to continue the computations according to pre-stored information.
+
+The education and research processes realized in this way allow getting acquainted with the theoretical basis and assist to develop the methods of design of the parallel algorithms aimed at solving time-consuming applied problems.
+More detailed information on the ParaLab system is presented in \cite{7_Gergel2010}.
+
+\section{The software system GlobalExpert for parallel solving the global optimization problems}
+
+GlobalExpert is an integrated software for solving the global optimization problems using the parallel computational architectures.
+This system has been developed on the base of the information-statistical theory of multiextremal optimization in the framework of which many efficient parallel global search algorithms have been designed [ ].
+
+Using GlobalExpert, the researchers can formulate the statement of the optimization problem, select an optimization algorithm from the list of the global search methods build-in in the system, set up the graphic indicators for monitoring the global search process, to execute the computations, obtain and analyze the results.
+
+GlobalExpert is oriented to the application by the engineers and researchers as an efficient tool for solving the multidimensional global optimization problems. 
+A considerable time of the computation of the optimized function at single point, which may reach tens of seconds is a key characteristic of the applied problems.
+In this case to solve optimization problems in a reasonable time GlobalExpert accumulates and analyzes scrupulously all optimization data computed in the search process.
+
+The system consists of two main parts: the optimization subsystem and the management one.
+The management subsystem is implemented as a .NET application using the managed C++ programming language.
+It is designed for the control of the computations and for displaying the calculated results.
+Using the management subsystem, the researchers get the following possibilities:
+
+\begin{itemize}
+\item \textit{to form an optimization problem} (to define the functions, to select the criteria and constraints, to define the feasible domain for the variables);
+
+\item \textit{to perform a preliminary investigation of the problem} (to plot the function sections and/or the level lines of the problem functions, to analyze the feasible domain of the problem fixing some variables);
+
+\item \textit{to select the optimization method} and to set its parameters;
+
+\item \textit{to select the running mode for the executive module} (serial, parallel on a single computer, parallel in a multiprocessor system);
+
+\item \textit{to monitor the search process in real time};
+
+\item \textit{to modify the problem statement and/or the method parameters} and to solve the problem in a new statement using the search information accumulated earlier.
+\end{itemize}
+
+A screenshot of the main window of the management subsystem is presented in Fig.~\ref{7_fig_3}).
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=1.0\linewidth]{figures/7_3.png}
+\caption{Modeling of the wheel profile}
+\label{7_fig_3}     
+\end{figure}
+
+The optimization subsystem is implemented in the form of an executable file as well as a module of a dynamic library (developed using С++ programming).
+In the form of the dynamic library module it can be linked to an application software system as a «solver» for the optimization problems.
+In the form of the executable file this module can be run either independently (the results are displayed in the console window) or under control of the management subsystem (the results are transferred to the management subsystem for visualizing and processing).
+
+When performing a large amount of computations the following procedure is recommended for solving time consuming optimization problems.
+First, one has to perform a preliminary calculations using the management subsystem, to ensure the correct problem statement and correctness of the results obtained when running the problem on a single computer or on a parallel computer with several processors.
+To run the computations on a real multiprocessor high-performance system, one has to run the executable file of GlobalExpert via the job management of the computational system.
+With this purpose, the generation of the command line with all options required for running a job on the parallel computer is provided in the management subsystem.
+
+Capacities of the system. GlobalExpert is an integrated program environment for solving the global optimization problems. It allows:
+
+\begin{itemize}
+\item \textit{to modify the statement of the problem} being solved without the loss of the accumulated search information (modification of the constraints set, change of the minimized function, expanding/reducing the search domain);
+
+\item \textit{to use a number of the information global search algorithms} (GlobalExpert includes several serial and parallel algorithms, which demonstrated the high efficiency performance in various applications);
+
+\item \textit{to display the visual information} on the problem being solved in the user-friendly form.
+The researcher can plot various sections of the problem functions, display the level lines of the minimized functions, the distribution of the iteration points on the function graphs and/or on the level lines, select the iteration points from different processors, etc.
+
+\item \textit{to accumulate and analyze the results} of the executed experiments.
+GlobalExpert provides logging of the executed experiments for storing the optimization results.
+The results of the computations include the search data from all processors, which the system has been run on.
+The accumulated data are presented in the tabular form, which is convenient for analysis and further processing.
+\end{itemize}
+
+\section{Optimization of the rail transport wheel profile}
+
+The considered problem of the search for the optimum profile of the wheels for the rail transport (railway, subway, tram, etc.) has been solved in the framework of joint research project supported by Russian Foundation for Basic Research (project number 04-01-89002-NVO\_a) and NWO (Netherlands Organization for Scientific Research, project number 047.016.014)  ``Fast Computing in Global Optimization: Sequential and Parallel Environments'', carried out at Lobachevsky State University of Nizhni Novgorod and Delft University of Technology, the Netherlands (TU Delft).
+
+\subsection{Problem Statement}
+
+The problem of the rail transport wheel profile optimization is described in details in \cite{8_Markine2005,8_Markine2007}.
+Here, a brief description of this problem is presented only.
+Since the wheels are of the conical shape, the center of a mounted axle is moving along a sinusoid.
+This process is illustrated in Fig.~\ref{8_fig_1}).
+The parameters of the contact between the wheel and the rail, such as the radius of rotation, the contact angle, and the angle of inclination for the mounted axle are varied at the transverse displacement of the mounted axle relative to the rail.
+The relationship between these variations and the transverse position of the mounted axle is determined by the wheel and rail profiles.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.9\linewidth]{figures/8_1.png}
+\caption{The displacement of the mounted axle}
+\label{8_fig_1}     
+\end{figure}
+
+The radius of rotation of the wheel at the contact point is an important parameter of the contact between the wheel and the rail.
+Actually, the radius can be different for the right and the left wheels since the mounted axle can shift relative to the rail (the radii $r_1$ and $r_2$, respectively in Fig.~\ref{8_fig_2}).
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.7\linewidth]{figures/8_2.png}
+\caption{The difference in the radii of rotation}
+\label{8_fig_2}     
+\end{figure}
+
+When the mounted axle is in the central position, the radii of rotation are the same for the left and right wheels, i.e., $r_1=r_2=r$.
+The difference in the radii of rotation for the left and right wheels can be defined as a function of the transverse displacement of the mounted axle relative to its central position $\Delta r(x)=r_1(x)-r_2(x)$.
+
+The mathematical model for this problem developed in Technical University Delft consists in the following.
+The wheel profile is described by B-spline, for building of which a set of points on the edge, on the edge base, and on the rolling surface of the wheel were selected (Fig.~\ref{8_fig_3})).
+Positions of these points can be varied for the purpose of variation of the profile.
+In order to reduce the computational costs for the optimization, the positions of the points on the upper surface of the edge as well as on the conical part of the profile were fixed since these parts of the wheel profile do not contact the rails.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.7\linewidth]{figures/8_3.png}
+\caption{Modeling of the wheel profile}
+\label{8_fig_3}     
+\end{figure}
+
+The ordinates of the movable points of the spline $z_i$ were selected as the components of the vector of the optimization problem parameters $y$, i.e.,
+\[
+y=(z_1,...,z_N),
+\]
+and the abscissas of these points were fixed.
+
+The number of the movable points and, consequently, the number $N$ of variables in the considered problem was equal to $11$.
+For each parameter   the interval $\left[–1, 1\right]$ was taken as the domain of its variation.
+The difference of the radii of rotation $\Delta r(x)$ was minimized.
+The constraints were introduced according to the stability considerations.
+For example, one of the constraints was imposed on the maximum angle of inclination for the mounted axle.
+
+\subsection{The results of the experiments}
+
+Thus, the optimization problem has $N=11$ parameters and $m=6$ constraints.
+The objective function as well as the constraint functions is the multiextremal ones.
+A one-dimensional section of the objective function is presented in Fig.~\ref{8_fig_4} for illustration.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.7\linewidth]{figures/8_4.png}
+\caption{A section of the objective function }
+\label{8_fig_4}     
+\end{figure}
+
+The constraints have no analytical form and defined by a MATLAB computational procedures.
+The computation of the values of all problem functions (objective function and constraints) at one point $y$ using Pentium IV processor with the clock frequency of 3 GHz takes about 10 sec.
+
+Thus, having formulated the problem, let us try to estimate the resources required for its solving in dependence on the method selected.
+Let us take, for example, the method of full scanning a uniform grid.
+Having selected 10 variants for the values of each parameter (remind, that the total number of parameters equals to 11), we obtain $10^{11}$ computations at the grid nodes that requires $\sim 10^{12}$ sec., i.e., $\sim 3$ years of computations using a supercomputer with 10 000 processors.
+Thus, the problem is so hard to solve that the parallel computations are not only necessary but should be supplied with the efficient numerical methods to obtain the results in a reasonable time unavoidably.
+
+The problem described above has been solved by the authors using the parallel index method with the multiple shifted evolvents on a cluster of 4 computers in Delft University of Technology.
+The search precision in the coordinates was $2^{-10}$ , i.e., 1024 points of each parameter value in the case of the full scanning method.
+The search time for the optimum estimate was 27 hours.
+The problem functions values have been computed $4297 + 4415 + 4236 + 4266 = 17214$ times.
+
+The computations carried out at TU Delft for a wheel of the optimum profile demonstrated its service life to increase up to 120 000 km between the profile corrections (more than five times longer as compared to the wheels of the original profile) and the maximum permitted speed to increase from 40 up to 60 m/sec.
+
+\section{Solving the inverse problem of chemical kinetics}
+
+In the present section the results of joint research project carried out at Lobachevsky State University of Nizhni Novgorod and at Institute of Petrochemistry and Catalysis, Russian Academy of Sciences (IPC RAS) \cite{8_Gubaidullin2011}.
+
+\subsection{Problem statement}
+
+The building of the mathematical models of the complex chemical reactions implies the presence of the unknown kinetic parameters (the rate constants, the activation energies, and the frequencies of collisions between the reacting molecules), which could be found by solving the problem of the minimization of the deviations between the calculated data (\textit{the direct problem}) and the experimental ones.
+Thus, a problem of identifying the mathematical model (\textit{the inverse problem of chemical kinetics}) arises, which in general case is a global optimization problem.
+
+Many problems of physics chemistry imply a considerable amount of computations, nevertheless, providing rather low precision.
+Investigating the inverse problems of chemical kinetics requires solving many systems of differential and algebraic equations.
+The kinetics of complex chemical reactions is featured by the presence of parameters varying fast and slowly (because various stages of reactions go with different rates).
+Therefore, solving the direct kinetic problems is complicated by the hardness of the differential equations describing the mechanisms of these reactions.
+
+The direct kinetic problem for the isothermal non-stationary model in a closed system is Cauchy problem for a system of ordinary differential equations
+\begin{equation} 
+ \frac{dx_i}{dt}= F_i, i=1,..,M; \; F_i=\sum_{j=1}^{N}{S_{ij}w_j}
+\end{equation}
+\begin{equation} 
+ w_j=k_j\prod_{i=1}^{M}{\left(x_i\right)^{\left|\alpha_{ij}\right|}}-k_{-j}\prod_{i=1}^{M}{\left(x_i\right)^{\left|\beta_{ij}\right|}}
+\end{equation}
+with the initial conditions $t=0$, $x_i(0)=x_i^0$, where 
+\begin{itemize}
+	\item $x_i$ are the concentrations of the substances (\textit{the molar fractions}) participating in the reaction;
+	\item $M$ is the number of substances; 
+	\item $N$ is the number of steps;
+	\item $S_{ij}$ are the stoichiometric matrix; 
+	\item $w_j$ are the rates of the $j$-th step, $1/{hr}$; 
+	\item $k_j$, $k_{-j}$ are the reduced rate constants for the forward and reverse reactions ($1/{hr}$), respectively;
+	\item $\alpha_{ij}$ are the negative elements of $S_{ij}$, and $\beta_{ij}$ are the positive elements of $S_{ij}$.
+\end{itemize}
+
+Since a part of the constants $k_j$, $k_{-j}$, as a rule, are not known, the problem of identification of the mathematical model of  the inverse kinetic problem arises.
+This problem is a minimization one for the function of the deviation between the calculated data and the experimental ones
+\begin{equation} \label{8_chem_func}
+F=\sum_{i=1}^n{\sum_{j=1}^M{\left|x_{ij}^p-x_{ij}^{\epsilon}\right|}}\rightarrow \min,
+\end{equation}
+where
+\begin{itemize}
+	\item $x_{ij}^p$ are the calculated values of the observable substances concentrations  (\textit{molar fractions});	
+	\item $x_{ij}^{\epsilon}$ are the values of the observable substances concentrations measured experimentally (\textit{molar fractions});
+	\item $n$ is the number of the experimental points.
+\end{itemize}
+
+To find the activation energies and the collision frequencies of the molecules reacting at the elementary step, Arrhenius equation
+\[
+k=Ae^{-\frac{Ea}{RT}}
+\]
+is used, where
+\begin{itemize}
+	\item $k$ is the reduced rate constant for the elementary step, $1/hr$;	
+	\item $E$ is the activation energy, $J/{mol}$;
+	\item $R$ is gas constant, $J/{(mol \cdot K)}$;
+	\item $А$ is the collision frequency for the reacting molecules;
+	\item $T$ is the temperature, $K$.	
+\end{itemize}
+
+Since in the search for the rate constants of the elementary steps, these ones may fall into the area, where the differential equations system describing the reactions may appear to be a stiff one, Mishelsen method with automatic step selection is applied to solving the direct problem \cite{8_Gubaidullin2010}.
+As the model output data $x_{ij}^p$ depend on the values of the constants $k_j$, $k_{-j}$ nonlinearly, problem (\ref{8_chem_func}) is a multiextremal optimization problem.
+
+\subsection{The results of the experiments}
+
+Traditionally, the methods based on the random search ideas had been applied in IPC RAS for solving problem (\ref{8_chem_func}).
+The typical computation costs were several days of continuous computations.
+Within the framework of the joint research project by IPC RAS and Lobachevsky State University of Nizhni Novgorod it was proposed to apply the efficient parallel global optimization methods developed in UNN.
+
+For the application of the novel identification technique the mathematical model for the reaction of the catalytical carboalumination of olefins and acetylenes with the assistance of threealkylalanes in the presence of transition metal complexes was selected.
+This reaction has applied in the laboratory research at IPC RAS as an efficient method for the synthesis of new Me-C (methyl-carbon), Et-C (ethyl-carbon), and C-C (carbon-carbon) bonds \cite{8_Parfenova2009}.
+The natural experiments of conducting the reactions of the catalytical carboalumination of olefins and acetylenes for different temperatures and fixed initial concentrations have been carried out in IPC RAS.
+In the experiments, the concentrations of five substances were traced and their percentage ratios were determined.
+A number of schemes for the mathematical description of the reaction of catalytical carboalumination of olefins and acetylenes (including the parameters to be identified) have been proposed in IPC RAS.
+As a priority problem of the identification of the unknown parameters, it was required to determine the rate constants for the elementary steps for the proposed reaction schemes for various temperatures as well as to find the activation energies for the elementary steps for various catalysts.
+These schemes and the identification problem have been described in details in \cite{8_Gubaidullin2011}.
+Here, note that the minimization problem arising depends on 10 parameters and is box-constrained one.
+The computational costs for the function value at one point are $\sim 3$ sec.
+(Intel Xeon 3.2 GHz).
+The objective function is  multiextremal that is confirmed by its two-dimensional section presented in Fig.~\ref{8_fig_5}.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.8\linewidth]{figures/8_5.png}
+\caption{Two-dimensional section of the objective function}
+\label{8_fig_5}     
+\end{figure}
+
+As a result of the computations, the estimates of the rate constants for the elementary steps for the 10- and 12-step reactions schemes of the carboalumination in the presence of $Cp_2 ZrCl_2$ catalyst have been obtained.
+The objective function value at the obtained optimizer was $F = 3.040$.
+It appeared to be impossible to achieve the function value close to zero because the experimental data contains the uncertainties.
+Nevertheless, the model identified using the parallel global optimization method allowed obtaining a good conformity between the calculated data and the experimental ones.
+As an illustration, the comparison of the computed and experimental concentrations of substances participating in the reactions is presented in Fig.~\ref{8_fig_6}.
+
+\begin{figure}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{figures/8_6a.png} \\ (a)}
+\end{minipage}
+\hfill
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{figures/8_6b.png} \\ (b)}
+\end{minipage}
+\caption{The comparison of the calculated data with the experimental ones}
+\label{8_fig_6}
+\end{figure}
+
+\section{Identification of the dynamic balance normative models of the regional economy}
+
+In the present section, the results obtained within the research project supported by RFBR ``Parallel global optimization methods in the identification of the dynamic balance for the normative models of the regional economy'', project 11-07-97017, carried out jointly by the research teams from Lobachevsky State University of Nizhni Novgorod and from Dorodnicyn Computing Centre of RAS are presented \cite{8_Gergel2011}.
+
+\subsection{Problem statement}
+
+The regional economy (the district, region, or state, etc.
+level) was the object of investigations.
+For the mathematical description of the economical processes taking place at the regional level, a dynamic balance normative economic model has been proposed in Dorodnicyn Computing Centre of RAS \cite{8_Olenev1999}.
+This model is a general one.
+However, it includes a significant number of parameters  being specific to the different regions.
+It is possible to find these parameters (or \textit{to identify the model}) by solving the problem of minimization of the deviation of the calculated time series of the macroeconomic measures from the corresponding historic statistical data.
+The functions entering the problem statement are the nonlinear ones.
+As a result, the problem of identification of the mathematical economic model is the global optimization one.
+
+Once the model was identified, it could be applied to verify various possible scenarios of the regional economic development on the base of the experiments with the model.
+The conclusions and predictions obtained using the identified mathematical model, in principle, could be disposed experimentally.
+This enables to correct further the model in order to obtain more justified prognosis based on the model being improved continuously.
+
+The problem of the identification of the multisector model of the regional economy presented here has been solved on the base of the data available for Nizhni Novgorod region, Russia.
+As it has been already mentioned above, the normative models include a large number of unknown parameters (the normatives of the distribution of the products and the normatives of distribution of the financial resources).
+In order to identify these parameters, it was proposed to apply the efficient parallel global optimization methods.
+The application of these algorithms on the supercomputer systems allows increasing the level of complexity of the mathematical models of the regional economy expressed by the number of the independent parameters.
+Carrying out the scenario computations based on the identified models of the regional economy allows more exact prognosis of the economic consequences of various strategic decisions.
+
+\subsection{The description of the model of the regional economy}
+
+The economy of Nizhni Novgorod region can be aggregated into three main sectors with approximately equal power:
+\begin{itemize}
+	\item the branches of the infrastructure, production and distribution of raw materials (agriculture, electric power industry, development, transportation, state administration, education, public health);
+	\item the manufacturing branches (machine building and metal working, chemical and petrochemical industry, fuel industry, ferrous and nonferrous metallurgy, construction materials industry, timber, woodworking, and pulp-and-paper industry, consumer goods industry, food, pharmaceutical, chemical, microbiological industry, flour-and-cereals and feed mill industry, printing industry, scientific complex);
+	\item the service branches (trading, operations with real estate, financial services, other services).
+\end{itemize}
+
+Let us select the following economic agents while designing the model: Regional Government, the producers represented by the three sectors selected above, the banking system, and the households of the region, the external vendors and consumers.
+
+In spite of absence of rich natural resources in Nizhni Novgorod Region its economy is one of the most developed one with respect to the industry among the regions of Russian Federation.
+The processing branches of industry, a powerful military industrial complex, well-developed fundamental scientific complex are the base of the industry in Nizhni Novgorod Region and there are good prospects for the development of the high-tech industries and of the innovative branches.
+
+The three-sector variant of the general balance model with supplies of products, factors of production, and financial resources in taxation and the illegal circulation of goods and finances \cite{8_Olenev2007} has been applied as the base for the design of the mathematical model of the regional economy.
+Remind that the producers in the model of Nizhni Novgorod Region economy are represented by three sectors:
+\begin{itemize}
+	\item the infrastructure complex ($X$),
+	\item the manufacturing branches complex ($Y$),
+	\item the services, mortgage, financial, and trading branches complex ($Z$).
+\end{itemize}
+The producers (the sectors of the regional economy $X$, $Y$, $Z$) use labor, capital, and the intermediate production of the partner sectors.
+The producers deliver their products to the internal markets and to the external ones.
+The households $L$ offer the labor and consume the final products.
+The resellers $T$ redistribute the goods and financial flows.
+The banking system $B$ lends money to the producers with the purpose to receive the banking profit.
+Regional Government $G$ collects the taxes from the producers and the households and manages the budget expenditures.
+Each market is assumed to establish its own prices for each kind of products.
+The variations of the prices are assumed to be inversely proportional to the variation of the supplies of the corresponding products.
+
+In order to account for the illegal circulation of the goods and money, the producers were assumed to divide the manufactured products into the legal part and the illegal one.
+The latter does not undergo the taxation.
+As a result, the producers receive two kinds of income: the ``legal'' and the ``illegal'' one.
+The ``illegal'' money can be laundered and the reserves undergo penalties.
+
+All the consumers’ money is assumed to be legal.
+The consumers’ income is assumed to be distributed between the acquisition of the legal products and the illegal ones from all sectors according to the predefined percentages.
+The production sectors $m = X,\;Y,\;Z$ pay the profit tax $n_1$, the value added tax $n_2$, the excise tax on the total output $n_3^m$ , the unified social tax on the payroll $n_4$, and the exports customs duties $n_5$.
+The households $L$ within the framework of the model pay the import customs duties $n_6$ and the income tax on the wages $n_7$.
+
+The criteria and the parameters of the model have the upper and lower indices.
+The upper indices are used for the agents and the lower ones for the valuables.
+The supplies of each valuable are distributed according to the normative  $a_i^{nm}$, which are the portions of the amount of the valuable $i$ being transferred from the agent $n$ to the agent $m$.
+The distribution of money is also performed according to some normative $b_i^{nm}$, which are the portions of the money resources of the agent $m$ transferred to the agent $n$ for the product $i$.
+The capital coefficients are also defined by the normatives $c_i^m$ , which are the normalized amount of the product $i$ spent for the producing a unit of a fund product for the agent $m$.
+
+The identification of the parameters was performed by the comparison of the output time series of the model variables with the available statistical time series in 2000--2008.
+As a criterion of proximity of the calculated time series to the statistical ones, Theil index was applied, which has been calculated according to the formula
+\[
+U=\frac{\sqrt{\sum{(x_i-y_i)^2}}}{\sqrt{\sum{x_i^2}}+\sqrt{\sum{y_i^2}}}
+\]
+The close the index to zero, the close the series compared.
+
+\subsection{The results of the experiments}
+
+The detailed description of the designed model is beyond the scope of this paper.
+The additional information can be found in \cite{8_Gergel2011}.
+Here we note that the global optimization problem arising in the model identification process has $60$ independent parameters, $86$ nonlinear constraints, and a multiextremal objective function.
+As an illustration, a two-dimensional section of the objective function confirming its essential multiextremality is presented in Fig.~\ref{8_fig_7}.
+
+\begin{figure}[t]
+%\sidecaption[t]
+\includegraphics[width=0.8\linewidth]{figures/8_7.png}
+\caption{Two-dimensional section of the objective function}
+\label{8_fig_7}     
+\end{figure}
+
+This problem has been solved on the cluster of Lobachevsky State University of Nizhni Novgorod using the parallel index method with the rotated evolvents (8 processors were employed).
+The estimate of optimum was obtained after $300 000$ iterations, the computation time of obtaining an estimate was $26$ minutes.
+Total computation time was $3$ hours.
+
+Using the identified mathematical model of Nizhni Novgorod Region, the scenario computations reflecting the dynamics of the economic indicators of the region for various economic policies (the pessimistic scenario, the basic, and the optimistic one) have been carried out in Computing Centre of RAS.
+
+\begin{thebibliography}{99.}
+
+\bibitem{7_Gergel1993} 
+Gergel, V.P.: A software system for multiextremal optimization. European Journal of Operational Research. \textbf{65(3)}, 305--313 (1993)
+
+\bibitem{7_Pinter1996}
+Pintér, J.D.: Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications). Dordrecht: Kluwer Academic Publishers (1996)
+
+\bibitem{7_Gergel2010}
+Gergel, V.P., Labutina, A.A.: The ParaLab system for investigating the parallel algorithms. Lecture Notes in Computer Science. \textbf{6083}, 95--104 (2010)
+
+\bibitem{8_Markine2005} 
+Shevtsov, I.Y., Markine, V.L., Esveld, C.: Optimal design of wheel profile for railway vehicles. Wear. \textbf{258(7-8)}, 1022--1030 (2005)
+
+\bibitem{8_Markine2007} 
+Shevtsov, I.Y., Markine, V.L., Esveld, C.: An inverse shape design method for railway wheel profiles. Structural and Multidisciplinary Optimization. \textbf{33(3)}, 243--253 (2007)
+
+\bibitem{8_Gubaidullin2011}
+Gubaidullin, I.M., Ryabov, V.V., Tikhonova, M.V.: Application of the global optimization index method to solving inverse problems of chemical kinetics.
+Numerical methods and programming.
+\textbf{12(1)}, 137--145 (2011) (in Russian)
+
+\bibitem{8_Gubaidullin2010}
+Tikhonova, M.V., Gubaidullin, I.M., Spivak, S.I.: The numerical solution of the direct chemical kinetics problem by the Rosenbrock's and Mishelsen's methods for the stiff systems of differential equations.
+MVMS Journal.
+\textbf{12(12}, 26--33 (2010) (in Russian)
+
+\bibitem{8_Parfenova2009}
+Parfenova, L.V., Gabdrakhmanov, V.Z., Khalilov, L.M., Dzhemilev, U.M.: On study of chemoselectivity of reaction of trialkylalanes with alkenes, catalyzed with $Zr$ $\pi$-complexes.
+Journal of Organometallic Chemistry.
+\textbf{694(23)}, 3725--3731 (2009)
+
+\bibitem{8_Gergel2011}
+Gergel, V.P., Gorbachev, V.A., Olenev, N.N., Ryabov, V.V., Sidorov, S.V.
+Parallel global optimization methods for identification of the dynamic balance normative model of regional economy.
+Bulletin of South Ural State University.
+\textbf{25(9)}, 4--15 (2011) (In Russian)
+
+\bibitem{8_Olenev1999}
+Avtukhovich, E.V., Guriev, S.M., Olenev, N.N., Petrov, A.A., Pospelov, I.G., Shananin, A.A., Chukanov, S.V.: A Mathematical Model of the Transition Economy.
+The Computing Centre of the Russian Academy of Sciences, Moscow (1999) (In Russian)
+
+\bibitem{8_Olenev2007}
+Olenev, N.N.: Model of innovation potential of regional economy.
+Proceedings of international conference ``Economy of depressed areas: problems and prospects of progress of regional economy''.
+Altai State University Press.
+178--188 (2007) (In Russian)
+
+\end{thebibliography}
 

### 2017-10-09 2017/URAL-PDC/u_pdc_paper.tex
u-pdc: fix some issues with English language

@@
 algorithm is based on the application of the information-statistical approach for solving
-reduced scalar problem, the set of the global optima in which coincides to the set of the
-weakly efficient solutions of the initial multi-objective problem. In the present work the parallelization scheme, which is common for all information-statistical global search algorithms,
+reduced scalar problem in which the set of the global optima coincides to the set of the
+weakly efficient solutions of the initial multi-objective problem.
+In the work the parallelization scheme, which is common for all information-statistical global search algorithms,
 was applied to the sequential algorithm of the multi-objective optimization. Also, in this
-work one of the techniques of accounting for the local properties of the objective function
-allowing accelerating the convergence essentially has been applied to the multi-objective
+work one of the convergence speedup techniques based on local properties of objective functions has been applied to the multi-objective
 problems.
@@
 The multi-objective optimization problems attract an increasing interest in recent years since
-these ones reflect the controversy of the requirements arising to the modeled objects of real
-world the most completely. In such problems, the solutions are a compromised decision of a set
+these ones most completely reflect the controversy of the requirements arising to the modeled objects of real
+world. In such problems, the solutions are a compromised decision of a set
 of these ones. As a rule, these compromised solutions are chosen either from a set, in which all
@@
 To find the complete set of efficient or weakly efficient points is the most computationally
-complex problem, however it provides a wide choice of possible compromised solutions. A
-plenty of approaches to the finding of these sets has been developed: the application of the
+complex problem, however it provides a wide choice of possible compromised solutions. To find these sets a
+plenty of approaches have been developed: the application of the
 genetic algorithms to the multi-objective problem directly~\cite{DebPratap2002}, the use of the
@@
 property in the case, when all criteria satisfy the Lipschitz condition has been developed on the
-basis of the information global search algorithm~\cite{markinStrongin1993}. In the present
-paper, the parallel and accelerated by the local refinement technique~\cite{strOptBook}(chapter
-3) modification of this method will be considered.
+basis of the information global search algorithm~\cite{markinStrongin1993}.
+In the paper there is considered the parallel modification of this method accelerated due to the local refinement
+technique~\cite{strOptBook}(chapter 3).
 
@@
 \end{equation}
-Let us assume the components of the vector function (the partial criteria) \(f_i(y), 1\leqslant
-i\leqslant m\), to satisfy the Lipschitz condition with the Lipschitz constant \(L_i\) in \(D\):
+Assume that the components of the vector function (the partial criteria) \(f_i(y), 1\leqslant
+i\leqslant m\) satisfy the Lipschitz condition with the Lipschitz constant \(L_i\) in \(D\):
 \begin{displaymath}
@@
 \end{equation}
-which is usually referred as the set of semi-efficient (or weakly efficient) solutions. The
+which is called the set of semi-efficient (or weakly efficient) solutions. The
 conditions in the right-hand side of the definition (\ref{eq:slater}) are known as the principle of
@@
 is a classic dimensionality reduction scheme for global optimization algorithms~\cite{evolvents2013}.
-Such a mapping allows the reduction of the problem (\(\ref{eq:problem}\)) stated in a
+Such a mapping allows to reduce the problem (\(\ref{eq:problem}\)) stated in a
 multidimensional space to solving a one-dimensional problem at the expense of worsening its
@@
 \par
-Therefore, not limiting the generality, one can consider the solving of the
+Without losing generality one can, one can consider the solving of the
 one-dimensional problem \(\min\{f(y(x)): x\in [0,1]\}\), satisfying the H\"older condition. The
@@
 have been considered in detail in~\cite{evolvents2013}. Here we would note that an evolvent
-built numerically is an approximation of the theoretical Peano curve with a precision of the
+built numerically approximates the theoretical Peano curve with a precision of the
 order \(2^{-m}\), where \(m\) is the building parameter of the evolvent.
@@
 \label{sec:test_problems}
-In order to evaluate the degree of speedup of the convergence of the modified algorithm from
+To evaluate the degree of speedup of the convergence of the modified algorithm from
 Section \ref{sec:algorithm}, the following problems were used:
@@
 Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM. The CPUs had 8 cores (i. e. total 16 cores
-were available per a node).
-In the present section, we will understand the speedup of the method as the speedup in the
+were available per a node). C++ language and OpenMP parallel framework were used to implement the considered algorithm.
+
+In this section, we will understand the speedup of the method as the speedup in the
 number of executed iterations, not in the time of execution. If the problem
@@
 
-\paragraph{Local refinement advantages.} In \cite{barkalovLebedev2016} the numerical
-experiments demonstrating the speedup of convergence when using the local refinement
-technique in the algorithm solving the scalar global optimization problems are presented. One
-could expect similar results from the multi-objective algorithm as well. The multi-objective
+\paragraph{Local refinement advantages.} In \cite{barkalovLebedev2016} the numerical experiments
+are presented which demonstrate the speedup of convergence of the scalar global optimization algorithm using the local refinement techniques described above.
+One could expect similar results from the multi-objective algorithm as well. The multi-objective
 algorithm with local refinement (MOAR) has been applied to the  Fonseca and Fleming
@@
 iterations decreased linearly with increasing value of \(p\) (Table
-\ref{tab:exResultsItersSpeedup}). In Fig. \ref{fig:s_estimation_all}, the examples of the
+\ref{tab:exResultsItersSpeedup}).
+Thus, the parallel algorithm reduces the number of iterations almost in \(p\) times comparing to the sequantal one.
+In Fig.
+\ref{fig:s_estimation_all}, the examples of the
 numerical solutions of the considered problems are given.
@@
 \section{Conclusion}
-In the present work, a method for solving the multi-objective optimization problems allowing
-finding a uniform estimate of the set of the weakly-optimal points has been considered. The
-techniques of parallelization and of the acceleration of convergence common for all algorithms
+In the work a method for solving the multi-objective optimization problem has been considered which allows to
+find a uniform estimate of the set of the weakly-optimal points.
+The techniques of parallelization and of the acceleration of convergence common for all algorithms
 of similar structure have been applied. The numerical experiments performed have

### 2017-10-15 2017/URAL-PDC/slides/slides.tex
u-pdc: add a placeholder for slides

\documentclass[aspectratio=1610]{beamer}

\usetheme{unnslides}
\usefonttheme{professionalfonts}

\usepackage{listings}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{cmbright}
\usepackage{fontspec}
\usepackage{unicode-math}
\usepackage{amsfonts}

\setromanfont{CMU Serif}
\setsansfont{CMU Sans Serif}
\setmathfont{Latin Modern Math}

\usepackage{polyglossia}
\setmainlanguage{russian}
%\setbeamertemplate{itemize item}{\color{black}$\blacktriangleright$}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\re}{Re}

\graphicspath{ {../paper/images/}{img/} }
%set pages numeration
\setbeamertemplate{footline}[frame number]
\setbeamertemplate{headline}{}
\setlength\abovecaptionskip{-1pt}

\title{Parallel Multi-objective Optimization Method for Finding Complete Set of Weakly Efficient Solutions}
\author{\textbf{Vladislav~Sovrasov}}
\institute{Lobachevsky University}
\date{}

\begin{document}
\begin{frame}[noframenumbering,plain]
\titlepage
\end{frame}

\begin{frame}
  \frametitle{Problem statement}

\end{frame}

\begin{frame}
  \frametitle{Dimension reduction}

\end{frame}

\begin{frame}
  \frametitle{Optimization method}

\end{frame}

\begin{frame}
  \frametitle{Test problems}
\end{frame}

\begin{frame}{{}}
  \frametitle{ }
  \begin{center}
    \Large{Q\&A}

\vspace{1cm}
    Vladislav Sovrasov

    sovrasov.vlad@gmail.com
  \end{center}
\end{frame}

\end{document}

### 2017-10-16 2017/URAL-PDC/slides/slides.tex
u-pdc: update slides

@@
 \usepackage{polyglossia}
-\setmainlanguage{russian}
 %\setbeamertemplate{itemize item}{\color{black}$\blacktriangleright$}
@@
   \frametitle{Problem statement}
-
+  \begin{displaymath}
+    \label{eq:problem}
+    \min\{f(y): y\in D\}, D=\{y\in \mathbb{R}^n: a_i \leqslant y_i \leqslant b_i, 1\leqslant i
+  \leqslant n \}.
+\end{displaymath}
 \end{frame}
@@
   \frametitle{Dimension reduction}
+  Peano-type curve \(y(x)\) allows to reduce dimension of the original multi-objective problem:
+  \begin{displaymath}
+  \lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant
+  N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
+  \end{displaymath}
 
+  \begin{eqnarray*}
+    \varphi(x)=\max\{h(x,y):y\in [0,1]\},x\in [0,1],\\
+    h(x,y)=\max\{f_i(x)-f_i(y):1\leqslant i\leqslant m\}.
+  \end{eqnarray*}
+
+  \begin{displaymath}
+    \varphi^*=\min\{\varphi(x):x\in [0,1]\}.
+  \end{displaymath}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Test problems}
+  \frametitle{Results}
+
+  Number of iterations decreases when increasing number of threads \(p\).
+  \begin{table}
+    \centering
+    \begin{tabular}{|l|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
+  \hline
+  \textbf{Problem} & \multicolumn{5}{c|}{\(p\)}\\
+  \cline{2-6}
+    & \(1\) & \(2\) & \(4\) & \(8\) & \(16\)\\
+  \hline
+  Markin-Strongin & 1041(198) & 516(198) & 256(185) & 131(197) & 68(191) \\
+  \hline
+  Fonseca and Fleming 2d & 1181(93) & 636(99) & 386(111) & 176(95) & 106(97) \\
+  \hline
+  Fonseca and Fleming 3d & 5346(160) & 3551(183) & 1186(143) & 606(153) & 351(142) \\
+  \hline
+  Viennet problem & 4896(276) & 2156(273) & 1226(270) & 631(287) & 286(274)\\
+  \hline
+  Poloni's function & 3351(102) & 1706(90) & 856(88) & 426(96) & 201(99) \\
+  \hline
+  \end{tabular}
+  \end{table}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results (speedup in iterations)}
+  \begin{table}
+    \centering
+    \begin{tabular}{|l|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
+  \hline
+  \textbf{Problem} & \multicolumn{4}{c|}{\(p\)}\\
+  \cline{2-5}
+    & \(2\) & \(4\) & \(8\) & \(16\)\\
+  \hline
+  Markin-Strongin & 2.02 & 4.07 & 7.95 & 15.31 \\
+  \hline
+  Fonseca and Fleming 2d & 1.86 & 3.06 & 6.71 & 11.14 \\
+  \hline
+  Fonseca and Fleming 3d & 1.51 & 4.51 & 8.82 & 15.23 \\
+  \hline
+  Viennet problem & 2.27 & 3.99 & 7.76 & 17.12\\
+  \hline
+  Poloni's problem & 1.96 & 3.91 & 7.87 & 16.67 \\
+  \hline
+  \end{tabular}
+  \end{table}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results (speedup in time)}
+  \begin{table}[ht]
+    \centering
+    \begin{tabular}{|l|p{1.6cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
+  \hline
+  \textbf{Problem} & \multicolumn{5}{c|}{\(p\)}\\
+  \cline{2-6}
+  &\(1\)(time, s) & \(2\) & \(4\) & \(8\) & \(16\)\\
+  \hline
+  Markin-Strongin & 104.47 & 1.97 & 3.65 & 6.79 & 9.90 \\
+  \hline
+  Fonseca and Fleming 2d & 118.95 & 1.85 & 2.81 & 5.79 & 6.40 \\
+  \hline
+  Fonseca and Fleming 3d & 554.45 & 1.51 & 4.14 & 8.05 & 10.69 \\
+  \hline
+  Viennet problem & 1488.6 & 2.22 & 3.64 & 6.98 & 13.49\\
+  \hline
+  Poloni's problem & 336.74 & 1.82 & 3.64 & 6.98 & 10.60 \\
+  \hline
+  \end{tabular}
+  \end{table}
 \end{frame}
@@
     sovrasov.vlad@gmail.com
+
+    https://github.com/sovrasov
   \end{center}

### 2017-10-17 2017/ParOptBook/Chapter4.tex
Chapters 4 and 5 are updated

@@
 \title{Global search algorithms for one-dimensional multiextremal optimization problems with nonlinear constraints}
-% Use \titlerunning{Short Title} for an abbreviated version of
-% your contribution title if the original one is too long
-\author{Name of First Author and Name of Second Author}
-% Use \authorrunning{Short Title} for an abbreviated version of
-% your contribution title if the original one is too long
-\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
-\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
+\titlerunning{Global optimization with nonlinear constraints}
+
+\author{Konstantin Barkalov, Victor Gergel, Roman Strongin}
+\authorrunning{K. Barkalov, V.Gergel, R. Strongin}
+
+\institute{Konstantin Barkalov,  Victor Gergel, Roman Strongin \at Lobachevsky State University of Nizhni Novgorod,  Nizhni Novgorod, Russia \email{konstantin.barkalov@itmm.unn.ru}}
 %

### 2017-10-17 2017/ParOptBook/Chapter5.tex
Chapters 4 and 5 are updated

@@
 %%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
+
 %\documentclass[graybox]{svmult}
@@
 
-As an illustration, let us consider an example of the maximization of a two-dimensional multiextremal function from the class (\ref{eq:3_518}) using the nested optimization scheme.
-The surface and level curves of the objective function are shown in Fig.
-5.1 and in Fig.
-\ref{fig:5_2}, respectively.
-The coordinates of the search trials of the nested scheme are shown in Fig.
-5.2 also.
+As an illustration, let us consider an example of using the nested optimization scheme for the maximization of a two-dimensional multiextremal function 
+\begin{equation}
+	\label{eq:5_GriFun}
+	\varphi(y)=\left\{\left(\sum _{i=1}^7\sum _{j=1}^7A_{ij} g_{ij} (y)+B_{ij}h_{ij}(y) \right)^2 +\left(\sum _{i=1}^7\sum _{j=1}^7C_{ij}g_{ij}(y)-D_{ij}h_{ij}(y)\right)^2 \right\}^{1/2},
+	\end{equation}
+	where 
+	\begin{equation*}
+	y=(y_1 ,y_2 )\in R^2 ,0\le y_s \le 1,s=1,2,
+	\end{equation*}
+	\begin{equation*}
+	g_{ij} (y)=\sin (i\pi y_1 )\sin (j\pi y_2 ),
+	\end{equation*}
+	\begin{equation*}
+	h_{ij} (y)=\cos (i\pi y_1 )\cos (j\pi y_2 )
+	\end{equation*}
+	 and the coefficients $A_{ij} ,B_{ij} ,C_{ij} ,D_{ij} $ are distributed uniformly over the interval $[-1,1]$.
+
+ The level curves of the objective function are shown  in Fig.
+\ref{fig:5_2}.
 For this function the global maximum value 13.51 is located at the point (0.603, 0.408).
-The method AGS (\ref{eq:2_38})--(\ref{eq:2_40}) was employed for solving the one-dimension subproblems (\ref{eq:5_21}).
+Fig.
+\ref{fig:5_2}contains also the coordinates of the search trials generated by the nested scheme with the information core global search algorithm  applied for solving the one-dimension subproblems (\ref{eq:5_21}).
 \begin{figure}[t]
@@
 which can be multidimensional with lesser dimensions.
-Such approach was developed (see \cite{5_BarkGer2014, 5_BarkGerLeb, 5_BarkLeb, 5_SysoyevBarkGerLeb}) in combination of the optimization algorithms based on the Peano space-filling curves for solving the nested multidimensional subproblems.
+Such approach was developed (see \cite{5_BarkGer2014, 5_BarkGerLeb, 5_BarkLeb, 5_SysoyevBarkGerLeb}) in combination with the optimization algorithms based on the Peano space-filling curves for solving the nested multidimensional subproblems.
 
@@
 The complexity of the second problem depends on the characteristics of the functions $\varphi^i(u_i)$ .
-These characteristics are influenced by the properties of the objective function $\varphi(y)$ as well as by the features of the search domain $Q$ defined, in turn, by constraints (\ref{eq:5_2}) and (\ref{eq:5_3}).
+These characteristics are influenced by the properties of the objective function $\varphi(y)$ as well as by the features of the search domain $Q$ defined by constraints (\ref{eq:5_2}) and (\ref{eq:5_3}).
 
@@
 
-Since the function $G(y)$ is assumed to be continuous in the domain $D$, the functions $G^i(u_i)$ are continuous with respect to $u_i\in D_i$  from (\ref{eq:5_11}) and, therefore, with regard to $y_i\in [a_i,b_i]$  as well.
-Then, at fixed $u_{i-1}\in D_{i-1}$  each one-dimensional problem (\ref{eq:5_21}) can be rewritten in a unified form
+Since the function $G(y)$ is assumed to be continuous in the domain $D$, the functions $G^i(u_i)$ are continuous in $D_i$  from (\ref{eq:5_11}) and, therefore, with regard to $y_i\in [a_i,b_i]$  as well.
+Then, with fixed $u_{i-1}\in D_{i-1}$  each one-dimensional problem (\ref{eq:5_21}) can be rewritten in a unified form
 \begin{equation}
@@
 
-The continuity of the constraint $g(x)$ allows one to state that the feasible domain $\bar{Q}$  can be presented as a system of intervals
+The continuity of the constraint $g(x)$ allows one to state that the feasible domain $\bar{Q}$  can be presented as a system of closed intervals
 \begin{equation}
@@
 \end{equation}
-in each of which the function $g(x)$ is non-positive.
+where in each interval the function $g(x)$ is non-positive.
 
@@
 \ref{fig:5_3}.
-The corresponding intervals  are marked with  the signs $\times$  on the abscissa axis including the points of tangency.
-The most right-side interval marked with $\times$ corresponds to the case when the function equals to zero over the interval of non-zero length.
+The corresponding intervals  are marked with  the signs $\times$  on the abscissa axis including the point of tangency.
+The right-side interval marked with $\times$ corresponds to the case when the function equals to zero over the interval of non-zero length.
 \begin{figure}[t]
@@
 \end{figure}
-In the system (\ref{eq:5_23}) the number of intervals $q$ can be infinite. As an example of such situation one can consider within the closed interval $[0,1]$ the function
+In the system (\ref{eq:5_23}) the number of intervals $q$ can be infinite. As an example of such situation the function
 \begin{displaymath}
@@
 \end{displaymath}
-
-Thus, in the case of a continuous function $G(y)$, the projection $\Pi_i(u_{i-1}$ determined as  (\ref{eq:5_16}) is a set of the kind (\ref{eq:5_23}), i.e.,
+within the closed interval $[0,1]$ can be considered.
+
+Thus, in the case of a continuous function $G(y)$, the projection $\Pi_i(u_{i-1})$ determined as  (\ref{eq:5_16}) is a set of the kind (\ref{eq:5_23}), i.e.,
 \begin{equation}
@@
 \end{equation}
-where, in general, the number of the intervals $q_i$  and their bounds $a_i^j$, $b_i^j$, $1\leq j\leq q_i,$ depend on the vector $u_{i-1}$ , i.e.,
+where, in general, the number of the intervals $q_i$  and their boundaries $a_i^j$, $b_i^j$, $1\leq j\leq q_i,$ depend on the vector $u_{i-1}$ , i.e.,
 \begin{equation}
@@
 
-If the domain $Q$  is such that it is possible to obtain the explicit (analytical) expressions for the values $q_i,a_i^j,b_i^j$   as  functions of $u_{i-1})\in Q_{i-1}$  for all $1\leq i\leq N$, then the domain $Q$ is called \textit{the domain with the computable bounds}.
+If the domain $Q$  is such that it is possible to obtain the explicit (analytical) expressions for the values $q_i,a_i^j,b_i^j$   as  functions of $u_{i-1})\in Q_{i-1}$  for all $1\leq i\leq N$, then the domain $Q$ is called \textit{the domain with the computable boundaries}.
 In particular, it is possible if all the roots of the functions $G^i(u_{i-1},y_i)$  with regard to the variables $y_i$  can be analytically found.
@@
 According to (\ref{eq:5_16}), the domains where the functions $G^1(y_1)$  for $y_1$  and $G^2(y_1,y_2)$  for $y_2$  are not positive determine the projections $\Pi_1$  and $\Pi_2(y_1)$ , respectively.
-The domain bounds are the roots of these functions in the interval $[0,2]$.
+The domain boundaries are the roots of these functions in the interval $[0,2]$.
 For the function $G^1(y_1)$  these roots are $0$ and $2$.
@@
 
-Thus, we can write down the bounds of projections (\ref{eq:5_24}) in an explicit form. Thereby, the domain $Q$ from (\ref{eq:5_26}) is a domain with computable bounds (\ref{eq:5_25}).
+Thus, we can write down the boundaries of projections (\ref{eq:5_24}) in an explicit form. Thereby, the domain $Q$ from (\ref{eq:5_26}) is a domain with computable boundaries (\ref{eq:5_25}).
 
@@
 
-In the example considered above, we have constructed the one-dimensional search domain bounds  analytically by finding the domains of the non-positivity of the corresponding functions $G^i(u_i)$ .
+In the example considered above, we have constructed the one-dimensional search domain boundaries  analytically by finding the domains of the non-positivity of the corresponding functions $G^i(u_i)$ .
 At the same time, one can specify a more visual «geometrical» method for the construction of the projections $\Pi_{i+1}(u_i)$.
 This method follows from definitions (\ref{eq:5_8}) and (\ref{eq:5_9}) and consists in the construction of the sections of the domain $Q$  by planes $u_i=const$.
-After that, it is necessary to find the bounds of these sections for the coordinate $y_{i+1}$.
+After that, it is necessary to find the boundaries of these sections for the coordinate $y_{i+1}$.
 
@@
 \end{displaymath}
-which roots of $g(x)$ are the global minima points in. To solve this problem, the characteristical global search algorithms providing the convergence to all global minimum points might be applied.
-
-Another approach to accounting the constraints in the optimization problems (the index method), which does not require to solve any auxiliary problems  is considered in Chapter 4.
+which roots of $g(x)$ are the global minima points in.
+To solve this problem, for example, the characteristical global search algorithms \cite{5_GrishaginSergeyevStrongin} providing the convergence to all global minimum points might be applied.
 
@@
 
-The optimization domains (\ref{eq:5_2}) can generate projections $\Pi_i(u_{i-1})$  as a single interval for some particular properties of the constraint functions as well.
-These properties are the convexity of the functions $g_j(y),\:1\leq j\leq m$, providing the convexity of the feasible domain $Q$ and the monotonous unimodality of the constraint functions \cite{5_GrishaginStrongin_EnginCybernetics}, which is more general than the concept of convexity.
+The optimization domains (\ref{eq:5_2}) can generate projections $\Pi_i(u_{i-1})$  as a single interval for some special classes of the constraint functions as well.
+These classes are the convex functions $g_j(y),\:1\leq j\leq m$, providing the convexity of the feasible domain $Q$ and the monotonous unimodal functions \cite{5_GrishaginStrongin_EnginCybernetics}, which is more general than the convex ones.
 In general case, monotonous unimodal constraints can generate non-convex domains $Q$.
@@
 \end{displaymath}
-the bounds of which, however, are not constant but depend on the vector $u_{i-1}$.
+the boundaries of which, however, are not constant but depend on the vector $u_{i-1}$.
 \subsection {Properties of the objective functions in one-dimensional subproblems}
@@
 
-The function $\varphi^2(y)\equiv \varphi(y)$ , obviously, is Lipschitzian with the constant $L$ with regard to the coordinate $y_2$.
+The function $\varphi^2(y)\equiv \varphi(y)$ , obviously, is Lipschitzian with the constant $L$ for the coordinate $y_2$.
 However, the function $\varphi^1(y_1)$  does not satisfy the condition (\ref{eq:5_29}).
@@
 \label{theor:5_1}
-Assume that the function $\varphi(y)$ satisfies the Lipschitz condition (\ref{eq:5_29})  with the constant $L>0$ in a domain $Q$ from (\ref{eq:5_2}) and the bound pairs (\ref{eq:5_27}) are the piecewise linear functions of the kind
+Assume that the function $\varphi(y)$ satisfies the Lipschitz condition (\ref{eq:5_29})  with the constant $L>0$ in a domain $Q$ from (\ref{eq:5_2}) and the boundary pairs (\ref{eq:5_25}) are the piecewise linear functions of the kind
 \begin{equation}
 \label{eq:5_30}
-a_{i+1}(u_i)=\max_{1\leq\nu \leq p_{i+1}}\{\alpha_i^\nu u_i+A_i^\nu\},
+a_i(u_{i-1})=\max_{1\leq\nu \leq p_i}\{\alpha_{i-1}^\nu u_{i-1}+A_{i-1}^\nu\},
 \end{equation}
@@
 \label{eq:5_31}
-b_{i+1}(u_i)=\max_{1\leq\nu \leq r_{i+1}}\{\beta_i^\nu u_i+B_i^\nu\},
-\end{equation}
-where  $\alpha_i^\nu u_i$ and $\beta_i^\nu u_i$ are the scalar products of the vectors from $R^i$  and $A_i^\nu, B_i^\nu$ are constants.
+b_i(u_{i-1})=\max_{1\leq\nu \leq r_i}\{\beta_{i-1}^\nu u_i+B_{i-1}^\nu\},
+\end{equation}
+where  $\alpha_{i-1}^\nu u_{i-1}$ and $\beta_{i-1}^\nu u_{i-1}$ are the scalar products of the vectors from $R^{i-1}$  and $A_{i-1}^\nu, B_{i-1}^\nu$ are constants.
 Then the functions $\varphi^i(u_i),u_i\in Q_i,$ satisfy the Lipschitz condition with the constants $L_i,1\leq i\leq N,$
@@
 
-Note that the representation of the boundy pairs in the form (\ref{eq:5_30}) and (\ref{eq:5_31}) takes place if the feasible domain is a convex polyhedron.
+Note that the representation of the boundary pairs in the form (\ref{eq:5_30}) and (\ref{eq:5_31}) takes place if the feasible domain is a convex polyhedron.
 
@@
 
-The main conclusion, which follows from the discussion of Lipschitz properties is that the properties of the objective functions of the one-dimensional problems depend essentially not only on the properties of the initial objective function $\varphi(y)$  but also on the shape of the feasible domain $Q$.
+The main conclusion, which follows from the discussion of the Lipschitz property is that the properties of the objective functions of the one-dimensional problems depend essentially not only on the features of the initial objective function $\varphi(y)$  but also on the shape of the feasible domain $Q$.
 
@@
 
-One of possible distributions for   $P=6$ parallel processors is presented in Fig. 
+One of possible distributions for   $P=6$ parallel processors is presented in Fig. \ref{fig:5_6}.
 \begin{figure}[t]
@@
 \begin{description} [1)]
-\item [1)] {AGSM– algorithm of global search which solves the one-dimensional subproblems inside of the nested scheme by the sequential core information algorithm AGS;}
-\item [2)] {PSIM – parallel synchronous information algorithm. It uses AGSIP for solving the one-dimensional subproblems (\ref{eq:5_21});}
-\item [3)] {AIPM – asynchronous multidimensional information parallel algorithm applying the asynchronous one-dimensional algorithm proposed in \cite{5_SergGriJCAA} for solving the one-dimensional subproblems.}
+\item [1)] {AGSM– algorithm of global search which solves the one-dimensional subproblems inside of the nested scheme by the sequential core information global search algorithm \cite{5_StrSergMon2000};}
+\item [2)] {PSIM – parallel synchronous information algorithm. It uses the algorithm AGSIP \cite{5_GriKvaMukhStr} for solving the one-dimensional subproblems (\ref{eq:5_21});}
+\item [3)] {AIPM – asynchronous multidimensional information parallel algorithm proposed in \cite{5_SergGriJCAA}.}
 \end{description}
@@
 
-This approach allows abstracting from the architecture of the parallel computational system, on which the experiments were carried out and estimate the «pure» efficiency of the algorithm decision rules.
+This approach allows abstracting from the architecture of the parallel computational system, on which the experiments are carried out and estimate the «pure» efficiency of the algorithm decision rules.
 
@@
 
-As a measure for the comparison of the relative efficiency of the synchronism and the asynchronism one can introduce the coefficient $u(\pi)=u_A(\pi)/u_S(\pi)$.
+As a measure for the comparison of the relative efficiency of the synchronism and the asynchronism the coefficient $u(\pi)=u_A(\pi)/u_S(\pi)$ is introduced.
 
@@
 
-In the first experiment, the maximization of 100 two-dimensional functions from the class (\ref{eq:3-58}) has been carried out.
+In the first experiment, the maximization of 100 two-dimensional functions from the class (\ref{eq:5_GriFun}) has been carried out.
 All experiments were carried out with the precision in the termination criterion $\epsilon=0.01$  and with the reliability parameter $r=2$  for solving all one-dimensional subproblems (\ref{eq:5_21}).
@@
 
-For the method AGSM, which is a pure sequential one, the average number of trials determined as a result of the maximization of 100 functions of the class (\ref{eq:3-58}) was found to be  $T_1= 415.08$.
+For the method AGSM, which is a pure sequential one, the average number of trials determined as a result of the maximization of 100 functions of the class (\ref{eq:5_GriFun}) was found to be  $T_1= 415.08$.
 This value can be considered as a measure for the average search time.
@@
 \end{table}
+
 The next series of the experiments consists in the minimization of the multidimensional function \cite{5_LucidiPiccioni}
@@
 
-Table 5.3 contains the results of the experiments A and B.
+Table \ref{tab:5_3} contains the results of the experiments A and B.
 \begin{table}
@@
 
-It results from Table \ref{tab:5_3} that both the methods make a small number of redundant trials only, and the asynchronous algorithm is faster than the synchronous one.
-
-Now let us consider a parallel characteristical algorithm.
-It is possible construct its generalization onto the synchronous and asynchronous cases like it has been done in Chapter 3 and compare the speed-ups of these versions as a measure of parallelization efficiency.
+It results from Table \ref{tab:5_3} that both the parallel methods make a small number of redundant trials only, and the asynchronous algorithm is faster than the synchronous one.
+
+Now let us consider a parallel characteristical algorithm \cite{5_GrishaginSergeyevStrongin}.
+It is possible construct its implementation for the synchronous and asynchronous cases and compare the speed-ups of these versions as a measure of parallelization efficiency.
 This approach allows estimating the potential of various parallelization schemes of the algorithm decision rules.
-In the framework of this approach, the experiments were carried out for core information global search algorithm (\ref{eq:2_36})--(\ref{eq:2_38}) for the  class of functions of various dimensionality.
-The average values of $u_S(\pi)$ and $u_A(\pi)$ obtained by the optimization of 100 multiextremal two-dimensional functions selected randomly from the test class (\ref{eq:2-58}) are presented in Fig.
+In the framework of this approach, the experiments were carried out for core information global search algorithm \cite{5_StrSergMon2000} for the  class of functions of various dimensionality.
+The average values of $u_S(\pi)$ and $u_A(\pi)$ obtained by the optimization of 100 multiextremal two-dimensional functions selected randomly from the test class (\ref{eq:5_GriFun}) are presented in Fig.
 \ref{fig:5_7}.
@@
 \centering
-%\includegraphics[width=0.8\linewidth]{figures/figure_5_7.pdf}
+\includegraphics[width=0.8\linewidth]{figures/figure_5_7.pdf}
 \caption{Comparative speed-ups of the synchronous scheme and of the asynchronous one on the two-dimensional test function class}
@@
 
-An analogous graph for the case of the optimization of a multiextremal five-dimensional function
+An analogous graph for the case of the optimization of the multiextremal five-dimensional function
 \begin{displaymath}
@@
 \ref{fig:5_8}.
-For the one-dimensional optimization the parallel version of the broken lines method (\ref{eq:2_36}), (\ref{eq:2_37}) with the adaptive estimate of Lipschitz constant (\ref{eq:2_39}), (\ref{eq:2_40}) was used.
+For the one-dimensional optimization the parallel version of the broken lines method (see \cite{5_GriKvaMukhStr, 5_Piyavskij}) with the adaptive estimate of Lipschitz constant was used.
 \begin{figure}[t]
 \centering
-%\includegraphics[width=0.8\linewidth]{figures/figure_5_8.pdf}
+\includegraphics[width=0.8\linewidth]{figures/figure_5_8.pdf}
 \caption{Speed-ups of the synchronous and asynchronous broken lines method}
@@
 
-As it follows from the presented structure the subproblems at the same level are solved independently and the search information obtained by a subproblem is not used in the other subproblems at this level (and, of course, in all sub-trees generated them).
+As it follows from the presented structure the subproblems at the same level are solved independently and the search information obtained in a subproblem is not used in the other subproblems at this level (and, of course, in all sub-trees generated them).
 Such loss of information leads to increasing the number of trials and slows up the optimization process.
@@
 To overcome this drawback in the paper \cite{5_GerGriGer} a new version of the nested optimization scheme in which the other way of solving the univariate subproblems has been proposed.
-In this version called adaptive nested optimization scheme instead of the strictly subordinated order of solving the univariate subproblems the principle of simultaneous consideration of all subproblems arising dynamically is used.
+In this version called \textit{adaptive nested optimization scheme} instead of the strictly subordinated order of solving the univariate subproblems the principle of simultaneous consideration of all subproblems arising dynamically is used.
 The adaptive scheme does not require to solve each  one-dimensional subproblem  up to the end and all the generated subproblems participate in the multidimensional optimization at any iteration of its realization.
-The main principle of such realization is close to the idea of characteristical algorithms.
+The main principle of such realization is close to the idea of characteristical algorithms  \cite{5_GriKvaMukhStr, 5_GrishaginSergeyevStrongin}.
 Namely, to each subproblem a numerical value called its characteristic is assigned and the multidimensional iteration consists in the choice of the subproblem with maximal characteristic and performing a new trial in this subtask.
-If the characteristics are assigned in such a way that they take into account the values of functions minimized in subproblems following the principle “low values influence increasing the characteristic” then the subproblems with lower values will be chosen more often than the subproblems with the higher ones and computational resources will not be spent for solving the latters unlike the classical nested scheme where these subproblems have to be solved completely.
+If the characteristics are assigned in such a way that they take into account the values of functions minimized in subproblems following the principle 'low values influence increasing the characteristic' then the subproblems with lower values will be chosen more often than the subproblems with the higher ones and computational resources will not be spent for solving the latters unlike the classical nested scheme where these subproblems have to be solved completely.
 
@@
 \begin{description} [\textbf{Step G1.}]
-\item [\textbf{Step G1.}] {Juxtapose to each subproblem from the set $S$ a numerical value called characteristic of this problem.}
+\item [\textbf{Step G1.}] {Juxtapose to each subproblem from the set $S$ a numerical value called \textit{characteristic} of this problem.}
 \item [\textbf{Step G2.}] {Choose the subproblem with the greatest characteristic (the currently best subproblem).}
@@
 A possible way can be suggested if for solving the subproblems (\ref{eq:5_21}) a one-dimensional characteristical algorithm is used.
-In this case as the characteristic of the subtask one can take the best interval characteristic (\ref{eq:2_31}) of this method.
-Moreover, as a stopping criterion of the multidimensional search the termination criterion (\ref{eq:2_21}) in the root problem (\ref{eq:5_19}) can be taken, i.e., the multidimensional optimization is stopped when the interval with the maximal characteristic in the subproblem (\ref{eq:5_19}) becomes less than predefined accuracy.
-
-For realization of this idea the core information  algorithm AGS (\ref{eq:2_37})--(\ref{eq:2_39}) can be taken.
+In this case as the characteristic of the subtask one can take the best interval characteristic of this method.
+Moreover, as a stopping criterion of the multidimensional search the termination criterion  in the root problem (\ref{eq:5_19}) can be taken, i.e., the multidimensional optimization is stopped when the interval with the maximal characteristic in the subproblem (\ref{eq:5_19}) becomes less than predefined accuracy.
+
+For realization of this idea the core information  global search algorithm  (see \cite{5_GriKvaMukhStr, 5_StrSergMon2000}) can be taken.
 Theorem ~\ref{theor:5_2} gives the theoretical substantiation of convergence to global minimum for the adaptive nested scheme combined with AGS for solving internal univariate subproblems (\ref{eq:5_21}).
@@
 \label{theor:5_2}
-Let the problem (\ref{eq:5_1})--(\ref{eq:5_3}) be considered without functional constraints (\ref{eq:5_2}), i.e., the admissible domain $Q$ coincides with the hyperparallelepiped $D$, and in the adaptive scheme  the information algorithm AGS (\ref{eq:2_37})--(\ref{eq:2_39}) be applied for one-dimensional optimization (\ref{eq:5_21}).
+Let the problem (\ref{eq:5_1})--(\ref{eq:5_3}) be considered without functional constraints (\ref{eq:5_2}), i.e., the admissible domain $Q$ coincides with the hyperparallelepiped $D$, and in the adaptive scheme  the information global search algorithm \cite{5_GriKvaMukhStr, 5_StrSergMon2000} be applied for one-dimensional optimization (\ref{eq:5_21}).
 If
@@
 \item {the objective function $\varphi(y)$  of the problem (\ref{eq:5_1}) satisfies in the domain $D$  the Lipschitz condition (\ref{eq:5_29}) with a finite constant $L>0$;}
-\item {during solving the subproblems (\ref{eq:5_21}) the parameter $m$ from (\ref{eq:2_39}) satisfies the inequality $m>2L$;}
-\item {while solving the root problem (\ref{eq:5_19}) the termination criterion (\ref{eq:2_21}) is used with $\epsilon=0$;}
-\item {for any function $\varphi^i(u_{i-1},y_i)$  of the family (\ref{eq:2_21}) the conditions 
+\item {during solving the subproblems (\ref{eq:5_21}) the parameter $m$ of the algorithm satisfies the inequality $m>2L$;}
+\item {while solving the root problem (\ref{eq:5_19}) the termination criterion is absent;}
+\item {for any function $\varphi^i(u_{i-1},y_i)$  of the family (\ref{eq:5_21}) the conditions 
 \begin{equation}
@@
 
-This theorem provides sufficient conditions of convergence to global minimum for the adaptive scheme in the combination with the method AGS.
+This theorem provides sufficient conditions of convergence to global minimum for the adaptive scheme in the combination with the core information global search algorithm.
 The proof of the theorem can be found in \cite{5_GerGriGer}.
@@
 \label{subsec:5_4_2}
-To demonstrate the efficiency of the adaptive nested scheme let us consider a function from the 2-dimensional test class (\ref{eq:3-58}) (see \cite{5_GrishaginOperChar, 5_StrSergMon2000}).
+To demonstrate the efficiency of the adaptive nested scheme let us consider a function from the 2-dimensional test class (\ref{eq:5_GriFun}).
 In Fig.~\ref{fig:5_10} the level curves of this function and distribution of trials (marked by points) are presented.
 The left panel corresponds to the adaptive scheme and the right panel to the classical one.
-For optimization of the function these schemes are used in combination with the method AGS for solving the univariate subproblems (\ref{eq:5_21}).
-In both the cases the method parameter $r=3$  and the search accuracy $\epsilon=0.01$ in termination criterion (\ref{eq:2_21}) were used.
+For optimization of the function these schemes are used in combination with the core information global search algorithm \cite{5_GriKvaMukhStr, 5_StrSergMon2000} for solving the univariate subproblems (\ref{eq:5_21}).
+In both the cases the algorithm parameter $r=3$  and the search accuracy $\epsilon=0.01$ in termination criterion were used.
 \begin{figure}[ht]
@@
 \item {10 local minima;}
-\item {standard function value $-1.0$ at the  global minimizer;}
+\item {standard function value -1.0 at the  global minimizer;}
 \item{distance 0.9 from the global minimizer to the vertex of the paraboloid;}
@@
 
-The following algorithms were compared by means of the method of operational characteristics – the adaptive nested scheme (AGS-A) and the classical nested scheme (AGS-C) with the univariate method AGS inside of both the schemes, and the very popular method DIRECT \cite{5_Jones, 5_JonesPerttunenStuckman} as an example of global optimization methods of different nature.
-Operational characteristics (see Subsection ~\ref{subsec:2_2_3} of Chapter 2) of the methods compared are presented in Fig. ~\ref{fig:5_11}.
+The following algorithms were compared by means of the method of operational characteristics: the adaptive nested scheme (AGS-A) and the classical nested scheme (AGS-C) with the core information global search algorithm inside of both the schemes, and the very popular method DIRECT \cite{5_Jones, 5_JonesPerttunenStuckman} as an example of global optimization methods of different nature.
+Operational characteristics (see \cite{5_GrishaginOperChar, 5_GriKvaMukhStr, 5_StrSergMon2000}) of the methods compared are presented in Fig. ~\ref{fig:5_11}.
 \begin{figure}[ht]
@@
 As well as the classical nested scheme, the adaptive one can be parallelized.
-The idea of parallelizing is similar to the approach applied in the parallel characteristical algorithms.
+The idea of parallelizing is similar to the approach applied in the parallel characteristical algorithms \cite{5_GrishaginSergeyevStrongin}.
 Namely, when in a multiprocessor environment there are $p>1$ free processors, in the decision rule of the general stage for the choice of the subproblems for further solving one can take $p$ subproblems with the best characteristics.
@@
 \bibitem{5_GriIsrCEUR}Grishagin, V.A., Israfilov, R.A.: Multidimensional Constrained Global Optimization in Domains with Computable Boundaries. CEUR Workshop Proceedings \textbf{1513}, 75--84 (2015)
+\bibitem{5_GriKvaMukhStr} Grishagin, V.A., Kvasov, D.E., Mukhametzhanov, M.S., Strongin, R.G.: Information and characteristical global search algorithms. This volume.
 \bibitem{5_GrishaginSergeyevStrongin} Grishagin, V.A., Sergeyev, Y.D., Strongin, R.G.: Parallel characteristical algorithms for solving problems of global optimization.

### 2017-10-17 2017/URAL-PDC/slides/slides.tex
u-pdc: add pictures for slides

@@
 \usepackage{amsfonts}
+\usepackage{subfig}
 
@@
 \begin{frame}
+  \begin{center}
   \frametitle{Problem statement}
+
   \begin{displaymath}
-    \label{eq:problem}
-    \min\{f(y): y\in D\}, D=\{y\in \mathbb{R}^n: a_i \leqslant y_i \leqslant b_i, 1\leqslant i
-  \leqslant n \}.
+    \min\{f(y): y\in D\}, D=\{y\in \mathbb{R}^n: a_i \leqslant y_i \leqslant b_i, 1\leqslant i \leqslant n \},
   \end{displaymath}
+where \(f(y)\) is a vector-function.
+
+\enspace
+Solution of the problem is a set of non-dominated points (Slater set):
+  \begin{displaymath}
+    S(D) = \{y\in D: \nexists z\in D, f_i(z)<f_i(y),1\leqslant i \leqslant m\}
+  \end{displaymath}
+Assume objectives to satisfy Lipschitz condition in \(D\):
+  \begin{displaymath}
+    |f_i(y_1)-f_i(y_2)|\leqslant L_i\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L_i<\infty,1\leqslant i\leqslant m
+  \end{displaymath}
+
+\end{center}
 \end{frame}
@@
   Peano-type curve \(y(x)\) allows to reduce dimension of the original multi-objective problem:
-  \begin{displaymath}
-  \lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant
-  N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
-  \end{displaymath}
+  \begin{gather}
+    \lbrace y\in \mathbb{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\} \nonumber \\
+    \min\{f(y): y\in D\}=\min\{f(y(x)): x\in [0,1]\} \nonumber
+  \end{gather}
 
+  \begin{figure}[ht]
+    \subfloat{{\includegraphics[width=.35\textwidth]{img/peano3d.png} }}
+    \subfloat{{\includegraphics[width=.35\textwidth]{img/peano2d.png} }}
+  \end{figure}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Scalarization tecnique}
+  Multi-objective problem can be reduced to a scalar optimization problem using the following function:
   \begin{eqnarray*}
-    \varphi(x)=\max\{h(x,y):y\in [0,1]\},x\in [0,1],\\
-    h(x,y)=\max\{f_i(x)-f_i(y):1\leqslant i\leqslant m\}.
+    h(x,y)=\min\{f_i(x)-f_i(y):1\leqslant i\leqslant m\}, \\
+    \varphi(x)=\max\{h(x,y):y\in [0,1]\},x\in [0,1].
   \end{eqnarray*}
-
+If \(x^*\) is weakly effective point, then \(\varphi(x^*)\le 0\), in the opposite case \(\varphi(x^*)> 0\), so
+the equivalent scalar problem is:
   \begin{displaymath}
@@
 \begin{frame}
+  \frametitle{Scalarization tecnique}
+  One-dimensional example:
+  \begin{figure}
+    \subfloat{{\includegraphics[width=.5\textwidth]{img/objectives.pdf} }}
+    \subfloat{{\includegraphics[width=.5\textwidth]{img/phi_func.pdf} }}
+  \end{figure}
+\end{frame}
+
+\begin{frame}
   \frametitle{Optimization method}
+Optimization method generates searh sequence \(\{x_k\}\) and consists of the following steps:
+\begin{enumerate}
+  \setlength{\itemindent}{.1in}
+  \item[Step 1.] Sort the searh information (one-dimentional points) in increasing order.
+  \item[Step 2.] Compute an approximation of the function \(\varphi(x)\).
+  \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
+  \item[Step 4.] Choose an interval \((x_{t-1}, x_t)\) with the greatest characteristic and
+  compute objective \(f(y(x))\) in the point choosen using the descision rule \(d\):
+  \begin{displaymath}
+    x^{k+1}=d(t)\in (x_{t-1}, x_t)
+  \end{displaymath}
+  \item[Step 4.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
+\end{enumerate}
+\end{frame}
 
+\begin{frame}
+  \frametitle{Improved parallel optimization method}
+  \begin{itemize}
+    \item
+    \textit{Local refinement.} Every \(q\) iterations ignore the characteristics \(R(i)\)
+    and perform calculation of the objective at the minimum point of the \(\varphi(x)\) approximation.
+    \item
+    \textit{Parallelization by characteristics.} At the Step 4 choose \(p\) best intervals
+    and generate \(p\) new points using rule \(d(i)\). Then compute the objective \(f(y(x))\) at this
+    point in parallel exploiting \(p\) computation units.
+
+    \enspace
+    In the best case (method doesn't generate redundand points compared to sequental
+    one and wastes time only on computation of the objective)
+    this scheme can give speedup in \(p\) times.
+  \end{itemize}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results}
+  Examples of numerical solutions:
+  \begin{figure}[ht]
+      \centering
+      \subfloat{{\includegraphics[width=.5\textwidth]{../img/viennet.png}}}
+      \subfloat{{\includegraphics[width=.5\textwidth]{../img/poloni.pdf} }}
+  \end{figure}
 \end{frame}

### 2017-10-18 2017/URAL-PDC/slides/slides.tex
u-pdc: update slides

@@
 \end{enumerate}
+\textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
 \end{frame}
@@
   \frametitle{Results}
+  The method was implemented on C++ language using OpenMP framework. All the experimants were carried out on
+  2xIntel Sandy Bridge E5-2660 2.2 GHz (total 16 cores) with 64Gb RAM.
+
   Examples of numerical solutions:
@@
 
-\begin{frame}
-  \frametitle{Results (speedup in iterations)}
-  \begin{table}
-    \centering
-    \begin{tabular}{|l|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
-  \hline
-  \textbf{Problem} & \multicolumn{4}{c|}{\(p\)}\\
-  \cline{2-5}
-    & \(2\) & \(4\) & \(8\) & \(16\)\\
-  \hline
-  Markin-Strongin & 2.02 & 4.07 & 7.95 & 15.31 \\
-  \hline
-  Fonseca and Fleming 2d & 1.86 & 3.06 & 6.71 & 11.14 \\
-  \hline
-  Fonseca and Fleming 3d & 1.51 & 4.51 & 8.82 & 15.23 \\
-  \hline
-  Viennet problem & 2.27 & 3.99 & 7.76 & 17.12\\
-  \hline
-  Poloni's problem & 1.96 & 3.91 & 7.87 & 16.67 \\
-  \hline
-  \end{tabular}
-  \end{table}
-
-\end{frame}
-
+%\begin{frame}
+%  \frametitle{Results (speedup in iterations)}
+%  \begin{table}
+%    \centering
+%    \begin{tabular}{|l|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
+%  \hline
+%  \textbf{Problem} & \multicolumn{4}{c|}{\(p\)}\\
+%  \cline{2-5}
+%    & \(2\) & \(4\) & \(8\) & \(16\)\\
+%  \hline
+%  Markin-Strongin & 2.02 & 4.07 & 7.95 & 15.31 \\
+%  \hline
+%  Fonseca and Fleming 2d & 1.86 & 3.06 & 6.71 & 11.14 \\
+%  \hline
+%  Fonseca and Fleming 3d & 1.51 & 4.51 & 8.82 & 15.23 \\
+%  \hline
+%  Viennet problem & 2.27 & 3.99 & 7.76 & 17.12\\
+%  \hline
+%  Poloni's problem & 1.96 & 3.91 & 7.87 & 16.67 \\
+%  \hline
+%  \end{tabular}
+%  \end{table}
+%
+%\end{frame}
+%
 \begin{frame}
   \frametitle{Results (speedup in time)}
+
   \begin{table}[ht]
@@
 
+\begin{frame}
+  \frametitle{Results (speedup and efficiency)}
+  \begin{figure}[ht]
+      \centering
+      \subfloat{{\includegraphics[width=.5\textwidth]{./img/speedup.pdf}}}
+      \subfloat{{\includegraphics[width=.5\textwidth]{./img/efficiency.pdf} }}
+  \end{figure}
+\end{frame}
+
 \begin{frame}{{}}

### 2017-10-18 2017/URAL-PDC/slides/draw_plots.py
u-pdc: add scripts for drawing pics

@@ -1,0 +1,46 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+def find_max(f, a, b, eps = 0.001):
+    n = int((b - a) / eps + 1 )
+
+    max_val = np.finfo(float).min
+
+    for i in range(n):
+        x = a + i*eps
+        max_val = max(max_val, f(x))
+
+    return max_val
+
+def main():
+
+    grid = np.linspace(-1, 3, 40)
+    pareto = np.linspace(0, 2, 20)
+    f1_values = np.power(grid, 2)
+    f2_values = np.power(grid - 2, 2)
+
+    plt.xlabel('$x$')
+    plt.plot(grid, f1_values, 'r--', label='$f_1(x)$')
+    plt.plot(grid, f2_values, 'g-', label='$f_2(x)$')
+    plt.plot(pareto, [0]*len(pareto), 'c-o', label='Slater set')
+    plt.grid()
+    plt.legend(loc = 'best', fontsize = 12)
+    plt.savefig('./img/objectives.pdf', format = 'pdf', dpi = 200)
+    plt.clf()
+
+    phi = lambda x: find_max(lambda y: min(x**2 - y**2, (x - 2)**2 - (y - 2)**2), -1, 3)
+    vec_phi = np.vectorize(phi)
+
+    plt.xlabel('$x$')
+    plt.plot(grid, f1_values, 'r--', label='$f_1(x)$')
+    plt.plot(grid, f2_values, 'g-', label='$f_2(x)$')
+    plt.plot(grid, vec_phi(grid), 'b-x', label='$\\varphi(x)$')
+    plt.grid()
+    plt.legend(loc = 'best', fontsize = 12)
+    plt.savefig('./img/phi_func.pdf', format = 'pdf', dpi = 200)
+
+if __name__ == '__main__':
+    main()

### 2017-10-18 2017/URAL-PDC/slides/draw_speedup_plots.py

@@ -1,0 +1,43 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+def main():
+    pRange = np.array([1, 2, 4, 8, 16])
+
+    s1 = (2.02 + 1.86 + 1.51 + 2.27 + 1.96) / 5
+    s2 = (4.07 + 3.06 + 4.51 + 3.99 + 3.91) / 5
+    s3 = (7.95 + 6.71 + 8.82 + 7.76 + 7.87) / 5
+    s4 = (15.31 + 11.14 + 15.23 + 17.12 + 16.67) / 5
+
+    sValues = np.array([1, s1, s2, s3, s4])
+
+    s1 = (1.97 + 1.85 + 1.51 + 2.22 + 1.82) / 5
+    s2 = (3.65+2.81+4.14+3.64+3.64) / 5
+    s3 = (6.79+5.79+8.05+6.98+6.98) / 5
+    s4 = (9.90+6.40+10.69+13.49+10.60) / 5
+
+    sTValues = np.array([1, s1, s2, s3, s4])
+
+    plt.xlabel('$p$')
+    plt.ylabel('Avg. speedup')
+    plt.plot(pRange[1:], sValues[1:], 'r-o', label = "Speedup in iterations")
+    plt.plot(pRange[1:], sTValues[1:], 'g-x', label = "Speedup in time")
+    plt.grid()
+    plt.legend(loc = 'best', fontsize = 12)
+    plt.savefig('./img/speedup.pdf', format = 'pdf', dpi = 200)
+    plt.clf()
+
+    plt.xlabel('$p$')
+    plt.ylabel('Avg. efficiency')
+    plt.plot(pRange, np.divide(sValues, pRange), 'r-o', label = "Efficiency in iterations")
+    plt.plot(pRange, np.divide(sTValues, pRange), 'g-x', label = "Efficiency in time")
+    plt.grid()
+    plt.legend(loc = 'best', fontsize = 12)
+    plt.savefig('./img/efficiency.pdf', format = 'pdf', dpi = 200)
+    plt.clf()
+
+if __name__ == '__main__':
+    main()

### 2017-10-19 2017/URAL-PDC/slides/slides.tex
u-pdc: add the final version of slides

@@
 \setromanfont{CMU Serif}
-\setsansfont{CMU Sans Serif}
+%\setsansfont{CMU Sans Serif}
 \setmathfont{Latin Modern Math}
@@
 \author{\textbf{Vladislav~Sovrasov}}
-\institute{Lobachevsky University}
+\institute{Lobachevsky University of Nizhni Novgorod}
 \date{}
@@
   \end{eqnarray*}
-If \(x^*\) is weakly effective point, then \(\varphi(x^*)\le 0\), in the opposite case \(\varphi(x^*)> 0\), so
+If \(x^*\) is weakly effective point, then \(\varphi(x^*)= 0\), in the opposite case \(\varphi(x^*)> 0\), so
 the equivalent scalar problem is:
@@
   \frametitle{Scalarization tecnique}
-  One-dimensional example:
+  Example:
   \begin{figure}
@@
   \end{displaymath}
-  \item[Step 4.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
+  \item[Step 5.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
 \end{enumerate}
@@
     \textit{Local refinement.} Every \(q\) iterations ignore the characteristics \(R(i)\)
-    and perform calculation of the objective at the minimum point of the \(\varphi(x)\) approximation.
+    and perform calculation of the objective near the current running-minimum point of the \(\varphi(x)\) approximation.
     \item
     \textit{Parallelization by characteristics.} At the Step 4 choose \(p\) best intervals
-    and generate \(p\) new points using rule \(d(i)\). Then compute the objective \(f(y(x))\) at this
-    point in parallel exploiting \(p\) computation units.
+    and generate \(p\) new points using rule \(d(i)\). Then compute the objective \(f(y(x))\) at these
+    points in parallel exploiting \(p\) computation units.
 
@@
   \frametitle{Results (speedup in time)}
-
+  Each calculation of the objective takes about 8ms.
   \begin{table}[ht]
@@
 
+\begin{frame}
+  \frametitle{Conclusion and future work}
+    Already done:
+    \begin{itemize}
+      \item local refinement technique was introduced to the multi-objective method scheme;
+      \item parallelism by characteristics was applied to the multi-objective method, efficiency metrics were collected.
+    \end{itemize}
+    Future work:
+    \begin{itemize}
+      \item compute quality metrics for the obtained solutions of the test problems;
+      \item compare the method with other approaches;
+      \item expand method to constrained problems;
+      \item use MPI for distributed computation of objectives.
+    \end{itemize}
+\end{frame}
+
 \begin{frame}{{}}
@@
 \vspace{1cm}
-    Vladislav Sovrasov
 

### 2017-10-25 2018/LION/LION2018.tex
short version (without superquantile)

--- 2017-04-12 2017/EUROGEN/eurogen.tex
@@
-% EUROGEN 2015 - Template for the extended abstract
-%defining format issues
-%**********************
+\documentclass{llncs}
 
-%Document Settings___________________________________________________________________
+\usepackage{hyperref}
+\usepackage{graphicx}
+\usepackage{multirow}
+\usepackage[misc,geometry]{ifsym}
 
-\documentclass[10pt,twocolumn,a4paper]{article}
-\special{papersize=210mm,297mm}
-
-%\makeatletter
-%\renewcommand\section{\@startsection{section}{1}{\z@}%
-%                                  {-3.5ex \@plus -1ex \@minus -.2ex}%
-%                                  {2.3ex \@plus.2ex}%
-%                                  {\normalfont\normalsize\bfseries}}
-%\makeatother
-
-%\makeatletter
-%\renewcommand\subsection{\@startsection{subsection}{1}{\z@}%
-%                                  {-3.5ex \@plus -1ex \@minus -.2ex}%
-%                                  {2.3ex \@plus.2ex}%
-%                                  {\normalfont\normalsize\em}}
-%\makeatother
-
-%________________________________________________________________________________
-
-%Miscellanneous__________________________________________________________________
-
-\sloppy % better line breaks
-\newcommand{\authorname}{\textbf}
-\newcommand{\authorgroup}{\emph}
-\newcommand{\authoraddress}{\emph}
-\newcommand{\authormail}{\emph}
-%________________________________________________________________________________
-
-%Packages________________________________________________________________________
-
-%\usepackage[latin1]{inputenc}
-\usepackage[utf8]{inputenx}
-\usepackage[compact]{titlesec}
-%\usepackage[pdftex]{graphicx}
-\usepackage[english]{babel}
-\usepackage{graphicx}
-\usepackage{amsmath,amssymb,mathrsfs}
-\usepackage{fancyhdr}
-\usepackage{calc}
-\usepackage{titling}
-\usepackage[T1]{fontenc}
-\usepackage{mathptmx}
-\usepackage[superscript]{cite}
-\usepackage{titlesec}
-\usepackage[none]{hyphenat}
+\usepackage{amsmath}
 \usepackage{enumitem}
-\usepackage{cite}
-%________________________________________________________________________________
-\titleformat*{\section}{\normalfont\normalsize\bfseries}
-\titleformat*{\subsection}{\normalfont\normalsize\em}
-\voffset = -1cm
-\textheight = 23.7cm
-\hoffset = -0.5cm
-\textwidth = 17cm
-\setlength\parindent{0.5cm}
-%Headers_________________________________________________________________________
-
-\pagestyle{fancy}
-\fancyhead[L]{EUROGEN 2017}\fancyhead[R]{September 13-15, 2017, Madrid, Spain}
-\fancyheadoffset[LE,RO]{\marginparsep + \marginparwidth}
-\setlength{\topmargin}{0pt}
-%________________________________________________________________________________
-
-%Maketitle Conf__________________________________________________________________
-\pretitle{\begin{center}\Large}
-\posttitle{\end{center}}
-
-\preauthor{\begin{center}\normalsize \lineskip 0.5em\begin{tabular}[t]{c}}
-\postauthor{\end{tabular}\par\end{center}}
-
-\predate{\begin{center}\tiny}
-\postdate{\par\end{center}}
-%________________________________________________________________________________
-%--------------------------------------------------------------------------------
-
-%Article data entry_____________________________________________________________
-
-\title{\vspace{-5ex}\textbf{Multipoint Approximation Method for Large Scale Design Optimization Problems Under Uncertainty}}
-\author{
-  \authorname{Vassili V. Toropov*}\\
-  \authorgroup{Professor of Aerospace Engineering, School of Engineering and Materials Science}\\
-  \authoraddress{Queen Mary University of London, London, UK}\\
-  \authormail{Email: v.v.toropov@qmul.ac.uk}\\\\
-  \authorgroup{Senior Research Fellow, Institute of Information Technology, Mathematics and Mechanics}\\
-  \authoraddress{Lobachevsky University of Nizhny Novgorod, Russia}\\\\
-  %--------------------------
-  \authorname{Yury M. Korolev}\\
-  \authorgroup{Research Fellow, Institute of Mathematics and Image Computing}\\
-  \authoraddress{University of Lübeck, Germany}\\
-  \authormail{Email: yury.korolev@mic.uni-luebeck.de}\\\\
-  %--------------------------
-  \authorname{Victor P. Gergel, Konstantin A. Barkalov}\\
-  \authorgroup{Institute of Information Technology, Mathematics and Mechanics}\\
-  \authoraddress{Lobachevsky University of Nizhny Novgorod, Russia}
-  %}
-}
-\date{}
-
-%_______________________________________________________________________________
-
-%Article Body___________________________________________________________________
+%\usepackage{amsfonts}
+%\usepackage{amssymb}
+%\usepackage{epstopdf}
+%\usepackage{epsfig}
 
@@
 
-%Abstract goes here--------------------------------------------------------------
+\mainmatter
 
-\twocolumn[
+\title{Multipoint Approximation Method for Large Scale Design Optimization Problems Under Uncertainty}
+\author{Vassili Toropov \inst{1,3} \and Yury Korolev \inst{2} \and Victor Gergel \inst{3} \and Konstantin Barkalov \inst{3} %\Letter 
+\\
+\email{konstantin.barkalov@itmm.unn.ru}}
+
+\institute{
+Queen Mary University of London, London, UK
+\and
+University of L\"ubeck, Germany
+\and
+Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
+}
 
@@
 
-\thispagestyle{fancy} %Insert header in title page
+\begin{abstract}
+The paper presents a new development in the Multipoint Approximation Method (MAM) that makes it capable of handling problems with uncertainty in design variables as well as in additional ‘environmental’ variables.
+The approach relies on approximations built in the combined space of design variables and environmental variables, and subsequent application of a risk measure and optimization with respect to the deterministic design variables, all within the iterative trust-region-based framework of MAM.
 
-\begin{center}
-\line(1,0){465}
-\end{center}
+\keywords global optimization, design optimization, multidisciplinary optimization, uncertainty, multipoint approximation method.
 
-\begin{@twocolumnfalse}
-
-\textbf{Summary}\\
-
-TBD\\
-
-\emph{\textbf{Keywords:} TBD}
-\begin{center}
-\line(1,0){465}
-\end{center}
-
-\end{@twocolumnfalse}
-
-]
-
-%--------------------------------------------------------------------------------
+\end{abstract}
 
@@
 
-Optimisation under uncertainty deals with responses $F(\pmb x,\pmb y)$ that depend on deterministic design variables $\pmb x$ and random ‘environmental’ variables $\pmb y$.
+Optimization under uncertainty deals with responses $F(\pmb x,\pmb y)$ that depend on deterministic design variables $\pmb x$ and random ‘environmental’ variables $\pmb y$.
 Uncertainty in the design variables can also be modelled by introducing additional random variables, e.g.
@@
 Randomness in $\pmb u$ and $\pmb y$ induces randomness in the responses $F(\pmb x+\pmb u,\pmb y)$.
-To convert them into deterministic functions, that can be optimised, a mapping $R$, called the \textit{risk measure}\cite{RockafellarUryasev2000} or \textit{robustness measure}\cite{SchillingsSchulz2015}, is applied to the responses.
+To convert them into deterministic functions, that can be optimized, a mapping $R$, called the \textit{risk measure} \cite{RockafellarUryasev2000} or \textit{robustness measure} \cite{SchillingsSchulz2015}, is applied to the responses.
 The result is then a deterministic function of the design variables: $\widetilde{F}(\pmb x)=R(F(\pmb x+\pmb u, \pmb y))$.
@@
 
-In problems with a large (in the order of hundreds) number of design variables, the multipoint approximation method (MAM\cite{Toropov1989,Toropov1992,ToropovFilatov1993}) proved to be efficient, e.g.
-in turbomachinery applications\cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010}.
+In problems with a large (in the order of hundreds) number of design variables, the multipoint approximation method (MAM \cite{Toropov1989,Toropov1992,ToropovFilatov1993}) proved to be efficient, e.g.
+in turbomachinery applications \cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010} .
 This method is an iterative optimization technique based on mid-range approximations built in trust regions.
@@
 The trust region will then translate and change size as optimization progresses.
-The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values\cite{KeulenToropovMarkine1996,ToropovKeulenMarkine1996} and occasional simulation failures\cite{ToropovMarkineHolden1999}.
-The mid-range approximations used in the trust regions, as originally suggested in Ref.
-\cite{Toropov1989} for structural optimization problems, are intrinsically linear functions (i.e.
+The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values \cite{KeulenToropovMarkine1996,ToropovKeulenMarkine1996} and occasional simulation failures \cite{ToropovMarkineHolden1999}.
+The mid-range approximations used in the trust regions, as originally suggested in \cite{Toropov1989} for structural optimization problems, are intrinsically linear functions (i.e.
 nonlinear functions that can be led to a linear form by a simple transformation) for individual sub-structures, and assembly of them for the whole structure.
-This was enhanced by the use of gradient-assisted metamodels,\cite{ToropovFilatov1993} use of simplified numerical models that is also termed a multi-fidelity approach,\cite{ToropovMarkine1996} and the use of analytical models derived by Genetic Programming\cite{ToropovAlvarez1998}.
-One of the recent developments\cite{PolynkinToropov2012} involved the use of approximation assemblies, i.e.
-a two stage approximation building process that is conceptually similar to the original one used in Ref.
-\cite{Toropov1989} but is free from the limitation that lower level approximations are linked to individual substructures.
+This was enhanced by the use of gradient-assisted metamodels \cite{ToropovFilatov1993}, use of simplified numerical models that is also termed a multi-fidelity approach,\cite{ToropovMarkine1996} and the use of analytical models derived by Genetic Programming \cite{ToropovAlvarez1998}.
+One of the recent developments \cite{PolynkinToropov2012} involved the use of approximation assemblies, i.e.
+a two stage approximation building process that is conceptually similar to the original one used in \cite{Toropov1989} but is free from the limitation that lower level approximations are linked to individual substructures.
 
-This paper presents a new development in the Multipoint Approximation Method that makes it capable of handling problems with uncertainty in design variables as well as in additional ‘environmental’ variables.
+The Moving Least Squares Method (MLSM) was proposed in \cite{LancasterSalkauskas1981} for smoothing and interpolation of scattered data and later used in the mesh-free form of the FEM \cite{Liszka1984}.
+As suggested in \cite{ChoiYounYang2001}, it can be used as a technique for metamodelling and used in MDO frameworks.
+The MLSM is a weighted least squares method where the weights depend on the Euclidean distance from a sample point to where the surrogate model is to be evaluated.
+The weight value for a certain sample point decays as the distance increases.
+Describing the weight decay with a Gaussian function tends to be the most useful option even though many others have been evaluated in \cite{ToropovSchrammSahaiJones2005}.
+As demonstrated in \cite{PolynkinToropov2010}, the cross-validated MLSM can be used both for design variable screening and for surrogate modelling.
+In order to create an efficient MDO framework for problems with disparate discipline attributes \cite{OllarToropovJones2014} extended the optimization approach of MAM to the use of local DOEs and MLS approximations built in different subspaces of the total design variable space corresponding to the individual disciplines.
+The subspaces are finally combined into the total design variable space in which the resulting MDO problem is solved.
+
+This paper presents a new development in the Multipoint Approximation Method that makes it capable of handling problems with uncertainty in design variables as well as in additional ``environmental'' variables.
 The approach relies on approximations built in the combined space of design variables and environmental variables, and subsequent application of a risk measure and optimization with respect to the deterministic design variables, all within the iterative trust-region-based framework of MAM.
@@
 Responses are evaluated at the DoE points and approximations are built using the obtained values.
-Currently, the pool of approximation methods available in MAM consists of a metamodel assemblies15 and the moving least-squares metamodels16-19.
+Currently, the pool of approximation methods available in MAM consists of a metamodel assemblies \cite{PolynkinToropov2012} and the moving least-squares metamodels \cite{LancasterSalkauskas1981,Liszka1984,ChoiYounYang2001,ToropovSchrammSahaiJones2005}.
 Other metamodel types could be used as well.
@@
     \end{equation}
-    The approximate problem (\ref{eq:problem_approx}) is solved using Sequential Quadratic Programming (SQP) and the solution of this problem   determines the centre of the next trust region.
-    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region10.
+    The approximate problem (\ref{eq:problem_approx}) is solved using Sequential Quadratic Programming (SQP) and the solution of this problem determines the center of the next trust region.
+    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region \cite{KeulenToropovMarkine1996}.
     \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $\pmb x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations).
@@
 
-This choice requires some changes in the MAM algorithm, since metamodels are now built in a larger space than the one where optimisation is performed in. The changes are summarised below.
+This choice requires some changes in the MAM algorithm, since metamodels are now built in a larger space than the one where optimization is performed in. The changes are summarized below.
 
@@
 \begin{enumerate}
-  \item Initialization: choose a starting point $\pmb x_0 \in {\Bbb R}^n$.
+%  \item Initialization: choose a starting point $\pmb x_0 \in {\Bbb R}^n$. 
+  \item Initialization: choose a starting point $\pmb x_0 \in {\pmb R}^n$. 
 	Choose an initial trust region $[\pmb a^0,\pmb b^0]$ in the combined space of design variables and environmental variables. 
 	Ensure that the initial guess $\pmb x_0$ belongs to the projection of the initial trust region onto the subspace of design variables.
-The projection of the initial trust region onto the subspace of environmental variables must cover the region where the realisations of the random variable y are likely to be.
-The coordinates of the trust region corresponding to the environmental variables will not be changed in the course of optimisation.
-Typically, the trust region along the coordinates corresponding to the $i$-th environmental variable is centred at the mean of the $i$-th environmental variable and has the width of six of its standard deviations.
+The projection of the initial trust region onto the subspace of environmental variables must cover the region where the realizations of the random variable y are likely to be.
+The coordinates of the trust region corresponding to the environmental variables will not be changed in the course of optimization.
+Typically, the trust region along the coordinates corresponding to the $i$-th environmental variable is centered at the mean of the $i$-th environmental variable and has the width of six of its standard deviations.
   \item At the $k$-th iteration the current approximation to the constrained minimum is $\pmb x^k$, the current trust region is $[\pmb a^k,\pmb b^k]$.
@@
 This distribution is assumed to be fixed.
-    The approximate problem (\ref{eq:problem_u_tilde}) is solved using Sequential Quadratic Programming (SQP) and the centre of the next trust region $\pmb x^{k+1}$ is determined as the solution of this problem.
-The optimisation runs in a subspace of the space where metamodels are built: metamodels are defined in the combined space of design variables and environmental variables, whilst optimisation runs in the subspace of the design variables.
+    The approximate problem (\ref{eq:problem_u_tilde}) is solved using Sequential Quadratic Programming (SQP) and the center of the next trust region $\pmb x^{k+1}$ is determined as the solution of this problem.
+The optimization runs in a subspace of the space where metamodels are built: metamodels are defined in the combined space of design variables and environmental variables, whilst optimization runs in the subspace of the design variables.
     \item The trust region is updated.
 The size of the trust region along the dimensions corresponding to the environmental variables is unchanged.
-The size of the next trust region along the dimensions corresponding to the design variables is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region according to the trust region strategy explained in Ref.
-\cite{KeulenToropovMarkine1996}.
+The size of the next trust region along the dimensions corresponding to the design variables is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region according to the trust region strategy explained in \cite{KeulenToropovMarkine1996}.
     \item The termination criterion is checked. If the termination criterion is satisfied, the algorithm proceeds to step 3. Otherwise, it returns to step 2
@@
 
-Summarising, the key changes introduced to account for uncertainties are (i) metamodels built in the combined space of design variables and environmental variables; (ii) optimisation performed in a subspace of the space where metamodels are defined; (iii) additive noise in the design variables; (iv) application of risk measures to transform random responses into deterministic functions.
+Summarizing, the key changes introduced to account for uncertainties are (i) metamodels built in the combined space of design variables and environmental variables; (ii) optimization performed in a subspace of the space where metamodels are defined; (iii) additive noise in the design variables; (iv) application of risk measures to transform random responses into deterministic functions.
 
@@
 As it has been already said, a risk measure is a mapping that transforms a random response $F(\pmb x+ \pmb u,\pmb y)$  into a deterministic function that can be optimized.
-Next we provide some examples of risk measures2,3.
+Next we provide some examples of risk measures \cite{RockafellarUryasev2000,SchillingsSchulz2015}.
 
@@
   \item $R(F)=\mu(F)$ where $\mu(F)$ is the mean of $F$.
-This choice corresponds to optimising average performance of the system.
+This choice corresponds to optimizing average performance of the system.
 In typical engineering applications, however, this is not sufficient and more conservative risk measures need to be chosen.
@@
 This risk measure is sometimes referred to as the safety margin.
-Optimisation using this risk measure is referred to as robust design optimisation22,23,24 and reflects the paradigms of a ‘$3\sigma$-design’, a ‘$6\sigma$-design’, etc.
+Optimization using this risk measure is referred to as robust design optimization \cite{KochYangGu2004,Yao2011,ZhangZhuChenArendt2013} and reflects the paradigms of a ‘$3\sigma$-design’, a ‘$6\sigma$-design’, etc.
   \item $R(F)=q_F(\alpha)$, where $q_F(\alpha)$ is the value of the quantile function, or simply the quantile, corresponding to a given probability $\alpha\in (0,1)$.
-This choice corresponds to the paradigm of reliability-based design optimization25,23,26.
+This choice corresponds to the paradigm of reliability-based design optimization  \cite{AouesChateauneuf2010,Yao2011,YounChoi2004}.
 In this approach, one wants to keep the probability of the system failing (i.e.
@@
 The value $1-\alpha$ is the desired probability of failure.
-Although rather intuitive, this choice has known computational issues which are discussed by Rockafellar and Royset27.
+Although rather intuitive, this choice has known computational issues which are discussed by Rockafellar and Royset \cite{RockafellarRoyset2010}.
 \end{itemize}
 
-For this study, we chose as risk measure the safety margin $R(F)=\mu(F)+k\sigma(F)$, aiming at covering both robust design optimisation and reliability design optimisation.
+For this study, we chose as risk measure the safety margin $R(F)=\mu(F)+k\sigma(F)$, aiming at covering both robust design optimization and reliability design optimization.
 Let us briefly discuss its computation.
@@
 To compute them efficiently, methods of high-dimensional integration must be employed.
-Here, quasi-Monte-Carlo integration30 based on Niederreiter sequences31 is used.
+Here, quasi-Monte-Carlo integration \cite{Caflisch1998} based on Niederreiter sequences \cite{Niederreiter1992} is used.
 
@@
 
-The example considered in this study is a classical engineering optimisation problem known as the scalable cantilevered beam34.
-The engineering object to be optimised is shown in Fig.
-\ref{fig:beam} taken from Ref.
-34.
+The example considered in this study is a classical engineering optimization problem known as the scalable cantilevered beam \cite{Vanderpllaats2001}.
+The engineering object to be optimized is shown in Fig.
+\ref{fig:beam} taken from \cite{Vanderpllaats2001}.
 
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{images/beam.png}
+    \includegraphics[width=1.0\textwidth]{beam.png}
     \caption{The cantilevered beam}
@@
 The design variables are the widths $b_i$ and heights $h_i$ of the segments.
-The number of segments $N$ can be chosen abitrary.
-The total length of the beam is 500 cm, the lenghths of the segments are $l_i=500/N$  cm.
+The number of segments $N$ can be chosen arbitrary.
+The total length of the beam is 500 cm, the lengths of the segments are $l_i=500/N$  cm.
 There are $N$ geometric constraints (the aspect ratios of each block (i.e.
@@
 The widths $b_i$ vary from 1.0 to 10.0 cm and the heights $h_i$ from 5.0 to 100.0 cm.
-The optimisation problem is formulated as follows:
+The optimization problem is formulated as follows:
 \begin{displaymath}
@@
 Hopefully, this explanation eliminates the confusion.
-Optimising $\mu(F) + k\sigma(F)$ with $k = 3.0$ using SQP on original functions, we obtain the value of the deterministic objective (volume of the structure) at the robust optimum $V = 67598.049\; cm^3$.
+Optimizing $\mu(F) + k\sigma(F)$ with $k = 3.0$ using SQP on original functions, we obtain the value of the deterministic objective (volume of the structure) at the robust optimum $V = 67598.049\; cm^3$.
 To compute $\mu(F)$ and $\sigma(F)$ at each iteration, 1000 points were used.
@@
 
-Recent developments in the Multipoint Approximation Method made it capable of solving large scale problems with uncertainty in parameters modelled as additional (‘environmental’) random variables and uncertainty in the design variables modelled as additive noise.
-The approach taken combines metamodels built in the combined space of design variables and environmental variables, application of risk measures that map random responses to deterministic functions, and optimisation in a subspace of the space where metamodels are defined, all within the iterative and trust-region-based framework of MAM.
-Performance is demonstrated on a benchmark example of structural optimisation known as the scalable cantilevered beam.
+Recent developments in the Multipoint Approximation Method made it capable of solving large scale problems with uncertainty in parameters modelled as additional (``environmental'') random variables and uncertainty in the design variables modelled as additive noise.
+The approach taken combines metamodels built in the combined space of design variables and environmental variables, application of risk measures that map random responses to deterministic functions, and optimization in a subspace of the space where metamodels are defined, all within the iterative and trust-region-based framework of MAM.
+Performance is demonstrated on a benchmark example of structural optimization known as the scalable cantilevered beam.
 
-\section*{Acknowledgement}
+\section*{Acknowledgment}
 The authors are grateful for the support provided by the Russian Science Foundation, project No. 16-11-10150.
@@
 \end{document}
-__________________________________________________________________________
+______________________________________________________________________

### 2017-10-25 2018/LION/LION2018.tex
Long version (with superquantile)

@@
 Although rather intuitive, this choice has known computational issues which are discussed by Rockafellar and Royset \cite{RockafellarRoyset2010}.
+  \item $R(F)=\overline{q}_F(\alpha)$, where $\overline{q}_F(\alpha)$ is the \textit{superquantile} introduced by Rockafellar and Uryasev \cite{RockafellarUryasev2000} and $\alpha \in (0,1)$ is a parameter.
+The $\alpha$-superquantile can be interpreted as the so-called \textit{buffered probability of failure}.
+The buffered probability of failure is always slightly more conservative than the usual probability of failure and, therefore, optimizing with respect to the buffered probability, one automatically gets constraints on the conventional probability of failure.
+It was found by Basova, Rockafellar, and Royset \cite{BasovaRockafellarRoyset} based on a number of engineering examples that the buffered probability of failure typically overestimates the conventional probability of failure by a factor of 3.
 \end{itemize}
 
-For this study, we chose as risk measure the safety margin $R(F)=\mu(F)+k\sigma(F)$, aiming at covering both robust design optimization and reliability design optimization.
+Superquantiles exhibit significantly better numerical properties than usual quantiles \cite{RockafellarRoyset2010}.
+The $\alpha$-superquantile can also be interpreted either as an average of quantiles  for $\alpha < \beta <1$, or as the conditional expectation of $F$ in its $(1−\alpha)$-tail \cite{RockafellarRoyset2015}:
+\[
+\overline{q}_F(\alpha) = \frac{1}{1-\alpha} \int_\alpha^1q_F(\beta)d\beta ;
+\]
+\[
+\overline{q}_F(\alpha) = E\left[F|F\geq q_F(\alpha)\right],
+\]
+where $E\left[F|F\geq q_F(\alpha)\right]$ is the conditional expectation of $F$ (conditional mean). The superquantile corresponding to $\alpha = 0$ is just the mean $\mu(F)=E[F]$.
+
+For this study, we chose two risk measures: the safety margin $R(F)=\mu(F)+k\sigma(F)$ and the superquantile $R(F)=\overline{q}_F(\alpha)$, aiming at covering both robust design optimization and reliability design optimization.
 Let us briefly discuss its computation.
@@
 Here, quasi-Monte-Carlo integration \cite{Caflisch1998} based on Niederreiter sequences \cite{Niederreiter1992} is used.
+
+Computation of the superquantile is not much more involved than calculating expectations, although it is not straightforward to see from its definition.
+A simple formula for the superquantile exists \cite{RockafellarUryasev2000,RockafellarRoyset2015} that combines computation of an expected value and solution of an auxiliary convex one-dimensional optimization problem:
+    \begin{equation} \label{eq:auxiliary_problem}
+			\overline{q}_F(\alpha) = \min_c\{c+E[(F-c)_+]\},
+    \end{equation}
+where where ($F-c)_+$ denotes the positive part of the function $(F(\pmb x+ \pmb u,\pmb y) - c)_+$ (i.e.
+$F(\pmb x+ \pmb u,\pmb y) - c)$, if this value is positive, and zero otherwise), $E[\cdot]$ is, as before, the expected value, and $c$ is a scalar.
+It has been shown by Rockafellar and Uryasev \cite{RockafellarUryasev2000} that the function to be minimized in (\ref{eq:auxiliary_problem}) is convex and continuously differentiable in $c$ if the underlying distribution is continuous.
+Therefore, the minimum in (\ref{eq:auxiliary_problem}) can be easily found using an appropriate one-dimensional convex optimization technique, e.g.
+the golden search, which was used in the examples in the next section.
+The integral (expectation) in (\ref{eq:auxiliary_problem}) can be approximated using the quasi-Monte-Carlo method.
+To solve the auxiliary optimization problem (\ref{eq:auxiliary_problem}), one would need to compute the expectation several times with different values of $c$.
+It should be noted, however, that this procedure does not require any additional evaluations of the response $F$ as the same samples can be used every time.
+Therefore, the solution of the auxiliary optimization problem (\ref{eq:auxiliary_problem}) does not present any significant additional computational challenges.
 
@@
 
+Finally, let us test the performance of MAM in optimization using the superquantile risk measure. The problem statement now is as follows:
+\begin{displaymath}
+  \begin{array}{c}
+    \min\limits_{\pmb b, \pmb h} \overline{q}_\alpha \left(V(\pmb b, \pmb h)\right) \\
+    s.t.\;1.0 \le b_i \le 10.0, \\
+    5.0 \le h_i \le 100.0, \\    
+		\overline{q}_\alpha \left(y_N\right) \le 2.5, \\
+    \overline{q}_\alpha \left(\sigma_i\right) \le \bar{\sigma} = 14000, \\
+    \overline{q}_\alpha \left(\frac{h_i}{b_i}\right)\le 20,
+  \end{array}
+\end{displaymath}
+where $\overline{q}_\alpha \left(F\right) $ denotes the $\alpha$-superquantile of a response $F$ (i.e.
+conditional expectation of $F$ in the upper $\alpha$-tail of its distribution).
+The value of the deterministic objective at the SQP solution in this problem for $\alpha = 1-10^{-3}$ (this corresponds to a probability of failure of $10^{-3}$) is $V = 68056.999\; cm^3$.
+
+Applying MAM to this problem, the value of the deterministic objective was obtained as $V = 68200.668\; cm^3$.
+Despite the fact that the superquantile uses information in the tail of the distribution, where local approximations used in MAM are not as reliable, the obtained results are very close to the reference SQP solution.
+The total number of  points used by MAM (i.e.
+the number of times the beam was analyzed) in this example is 1649.
+
 The $\mu(F ) + k\sigma(F)$ risk measure uses quantities that are estimated in the middle of the distribution --- the mean and the standard deviation.
@@
 On the other hand, probabilistic interpretation of the obtained results in this case is only possible under the normality assumption for the distribution of $F$; if the actual distribution of $F$ has heavier tails than the normal one, the actual probability of failure will be higher than the one  the system was designed for.
+
+The superquantile focuses on the tails of the distribution, and is better computed using importance sampling with a distribution with heavier tails (in this example, the Student's $t$-distribution was used).
+Although the superquantile somewhat overestimates the probability of failure (typically by the factor of 3, as reported by Basova, Rockafellar and Royset \cite{BasovaRockafellarRoyset} based on several of engineering examples), it provides a distribution-independent constraint on the probability of failure of an engineering system being optimized.
 

### 2017-10-31 2017/ParOptBook/Chapter1.tex
References and some misprints are corrected

@@
 \bibitem{1_GriIsrAIP}	Grishagin, V.A., Israfilov, R.A.: Global search acceleration in the nested optimization scheme. AIP Conf. Proc. \textbf{1738}, 400010 (2016)
-\bibitem{1_GriIsrSerg}	Grishagin, V.,  Israfilov, R., Sergeyev, Y.: Convergence conditions and numerical comparison of global optimization methods based on dimensionality reduction schemes.
+\bibitem{1_GriIsrSerg}	
+Grishagin, V.,  Israfilov, R., Sergeyev, Y.: Convergence conditions and numerical comparison of global optimization methods based on dimensionality reduction schemes.
 Appl.
@@
 Comput.
-(2017) doi:10.1016/j.amc.2017.06.036
+\textbf{318}, 270--280 (2018)
+
 \bibitem{1_GriIsrCEUR} Grishagin, V.A., Israfilov, R.A.: Multidimensional Constrained Global Optimization in Domains with Computable Boundaries. CEUR Workshop Proceedings 1513, 75--84 (2015)

### 2017-11-02 2018/LaTeX/Loc_eng.tex
Added a folder remotely

\documentclass[a4paper]{locconf}
\usepackage{graphicx}
\usepackage[russian,english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\begin{document}
\title{High Performance Computing for Global Optimization Problems}

\author{K.A. Barkalov, V.P. Gergel}

\address{Lobachevsky State University of Nizhni Novgorod, Gagarin avenue 23, Nizhni Novgorod, Russia, 603950}

\begin{abstract}
All articles must contain an abstract. The abstract should give readers concise information about the content of the article and indicate the main results obtained and conclusions drawn. The abstract is not part of the text and should be complete in itself; no table numbers, figure numbers, references or displayed mathematical expressions should be included. It should be suitable for direct inclusion in abstracting services and should not normally exceed 200 words in a single paragraph. Since contemporary information-retrieval systems rely heavily on the content of titles and abstracts to identify relevant articles in literature searches, great care should be taken in constructing both.
\end{abstract}

\section{Introduction}

This file should be used as a pattern for English Thesis. You can write on email mingazov88@gmail.com with questions concerning the preparation of the English Thesis.

\section{Problem statement}

Let us consider the $N$-dimensional global optimization problem
\begin{equation}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y): y \in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}
\end{equation}
with search domain
\begin{equation}\label{D}
D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.
%D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.
\end{equation}

The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and
the left-hand sides $g_i(y), \; 1\leq i \leq m$, of the constraints
satisfy the Lipschitz conditions with constants $L_i, \; 1 \leq i \leq
m+1$, respectively, and may be multiextremal. It is assumed that the
functions $g_i(y),\; 1 \leq i \leq m+1$, are defined and computable only in the corresponding domains
\[
Q_1=D, \; Q_{i+1}=\left\{y \in Q_i : g_i(y) \leq 0 \right\}, \; 1 \leq i \leq m.
\]

These conditions allow for the introduction of a classification of the
points $y \in D$ according to the number $\nu (y)$ of the constraints
computed at this point. 
%отсюда

Thus, a \textit{trial} at a point $y^k \in D$ executed at the $k$-th
iteration of the algorithm will consist in computing the values $g_1(y),...,g_\nu(y),$ where the
index $\nu \leq m$ is determined by the conditions
	\[
		g_i(y^k )\leq 0, \; 1 \leq i < \nu, \; g_\nu(y^k)>0, \; \nu \leq m.
	\]
The occurrence of the first violation of the constraint terminates the
trial. In the case when the point $y^k$ is a feasible one, i.e. when
$y \in Q_{m+1}$, the trial includes the computation of the values of
all the functions of the problems and the index is assumed to be
$\nu=m+1$. The pair of values
\[
\nu=\nu(y^k), \; z^k=g_\nu(y^k)
\]
is a \textit{trial result}.

\Russian
Параллельный индексный алгоритм, который может быть применен для решения подобных задач с частично вычислимыми функциями и который реализован в системе Globalizer, подробно описан в [2]. Основная идея алгоритма состоит в сведении исходной многомерной задачи к набору связанных подзадач меньшей размерности, и параллельному их решению. Схемы редукции размерности, используемые при работе индексного алгоритма, кратко описаны в следующем разделе.

\section*{Acknowledgments}
This study was supported by the Russian Science Foundation, project No 16-11-10150.

\section*{References}

\medskip

\begin{thebibliography}{9}

\bibitem{Strongin2000}
Strongin, R.G. Global optimization with non-convex constraints. Sequential and parallel algorithms. / R.G. Strongin, Ya.D.  Sergeyev --- Dordrecht: Kluwer Academic Publishers, 2000. --- 704 p.

\bibitem{Strongin2013}
Strongin, R.G. Parallel computing in global optimization problems / R.G. Strongin, V.P. Gergel, V.A. Grishagin, K.A. Barkalov --- Moscow: Publishing of the Moscow State University, 2013. --- 280 p. --- (in Russian).

\bibitem{Jones}
Jones, D.R. The DIRECT global optimization algorithm / In: Floudas, C. A., Pardalos, P. M. (eds.) The Encyclopedia of Optimization, Second Edition. --- Heidelberg: Springer, 2009. --- P. 725--735.

\bibitem{Zilinskas2011}
Paulavi\v{c}ius, R. Parallel branch and bound for global optimization with combination of Lipschitz bounds / R. Paulavi\v{c}ius, J. \v{Z}ilinskas, A. Grothey // Optimization Methods \& Software. --- 2011. --- Vol. 26(3). --- P. 487--498.

\bibitem{Evtushenko}
Evtushenko, Yu.G. Parallel global optimization of functions of several variables / Yu.G. Evtushenko, V.U. Malkova, A.A. Stanevichyus // Computational Mathematics and Mathematical Physics. --- 2009. --- Vol. 49 (2). --- P. 246--260.

\bibitem{Sysoyev}
Sysoyev, A.V. MPI implementation of dimension reduction multilevel scheme for parallel solving the global optimization problems / A.V. Sysoyev, K.A. Barkalov, V.P. Gergel, I.G. Lebedev // Russian Supercomputing Days: Proceedings of the International Scientific Conference. --- Moscow: Publishing of the Moscow State University, 2015. --- P. 61--68.

\bibitem{BarkalovLebedev}
Barkalov, K.A. Solving global optimization problems on GPU / K.A. Barkalov, I.G. Lebedev // Russian Supercomputing Days: Proceedings of the international conference. --- Moscow: Publishing of the Moscow State University, 2016. --- P. 640--650.

\bibitem{BarkalovGergel}
Barkalov, K. Parallel global optimization on GPU / K. Barkalov K., V. Gergel // Journal of Global Optimization. --- 2016. --- Vol. 66(1). --- P. 3--20.

\bibitem{BarkalovGergelLebedev}
Barkalov, K. Use of Xeon Phi Coprocessor for Solving Global Optimization Problems / K. Barkalov, V. Gergel, I. Lebedev //  Lecture Notes in Computer Science. --- 2015. --- Vol. 9251. --- P. 307--318.

\bibitem{Gergel2017}
Gergel, V. An Approach for Generating Test Problems of Constrained Global Optimization / V. Gergel // Lecture Notes in Computer Science. --- 2017. --- Vol. 10556. --- P. 314--319.

\bibitem{Gergel2016}
Gergel, V. Adaptive nested optimization scheme for multidimensional global search / V. Gergel, V. Grishagin, A. Gergel // Journal of Global Optimization. --- 2015. --- Vol. 66(1). --- P. 35--51.

\bibitem{Sergeyev2017}
Sergeyev, Ya.D. Deterministic global optimization: An introduction to the diagonal approach / Ya.D. Sergeyev, D.E. Kvasov --- New York: Springer, 2017. --- 136 p.

\end{thebibliography}

\end{document}

### 2017-11-03 2018/ITNT/ITNT2018.tex
Pictures ans tables

--- 2017-11-02 2018/LaTeX/Loc_eng.tex
@@
 \Russian
-Параллельный индексный алгоритм, который может быть применен для решения подобных задач с частично вычислимыми функциями и который реализован в системе Globalizer, подробно описан в [2].
-Основная идея алгоритма состоит в сведении исходной многомерной задачи к набору связанных подзадач меньшей размерности, и параллельному их решению.
-Схемы редукции размерности, используемые при работе индексного алгоритма, кратко описаны в следующем разделе.
+ѕараллельный индексный алгоритм, который может быть применен дл€ решени€ подобных задач с частично вычислимыми функци€ми и который реализован в системе Globalizer, подробно описан в [2].
+ќсновна€ иде€ алгоритма состоит в сведении исходной многомерной задачи к набору св€занных подзадач меньшей размерности, и параллельному их решению.
+—хемы редукции размерности, используемые при работе индексного алгоритма, кратко описаны в следующем разделе.
+
+\section{Dimensionality reduction}
+
+\subsection{Dimensionality reduction using space-filling curves}
+
+The use of Peano curve $y(x)$ 
+\begin{equation}
+\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq N\right\}=\left\{y(x):0\leq x \leq 1 \right\}
+\end{equation}
+unambiguously mapping the interval of real axis $[0,1]$ onto a $N$-dimensional cube is the first of the dimensionality reduction methods considered.
+To implement this method of dimensionality reduction a numerically constructed curve (\textit{evolvent}) is used.
+The evolvent is $2^{-m}$ accurate approximation of the theoretical Peano curve in $L_\infty$ metric, where $m$ is an evolvent construction parameter.
+Problems of numerical construction of the evolvents and the corresponding theory are considered in detail in \cite{Strongin2000}.
+
+By using this kind of mapping it is possible to reduce the multidimensional problem~(\ref{problem}) to a univariate problem
+\[
+\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
+\]
+
+The considered dimensionality reduction scheme juxtaposes a
+multidimensional problem with lipschitzian functions to a univariate
+problem where the corresponding functions satisfy the uniform H{\"o}lder
+condition (see \cite{Strongin2000}), i.e.
+\[
+\left|g_i(y(x'))-g_i (y(x''))\right| \leq H_i \left|x'-x'' \right|^{1/N}, \; x',x''\in [0,1], \; 1\leq i \leq m+1.
+\]
+Here $N$ is the dimensionality of the initial multidimensional problem and
+the coefficients $H_i$ are related with the Lipschitz constants $L_i$ of
+the initial problem by the inequalities $H_i \leq 2L_i \sqrt{N+3}$.
+
+\subsection{Nested optimization scheme}
+
+Nested optimization scheme is based on relation (see \cite{Grishagin2001})
+\begin{equation}\label{nested}
+\min_{y \in D}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}= \min_{u_1\in D_1}\min_{u_2\in D_2}...\min_{u_M\in D_M }{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
+\end{equation}
+which allows replacing the solving of multidimensional problem (\ref{problem}) by solving a family of subproblems related to each other recursively.
+Here we consider vector $y$ as a vector of block variables
+\begin{equation}
+y=(y_1,y_2,...,y_N)=(u_1,u_2,...,u_M),
+\end{equation}
+where the $i$-th block variable $u_i$ is a vector of $N_i$ components of vector $y$, taken serially i.e.
+$u_1=(y_1,y_2,...,y_{N_1})$, $u_2=(y_{N_1+1},y_{N_1+2},...,y_{N_1+N_2})$,..., $u_M=(y_{N-N_M+1},y_{N-N_M+2},...,y_{N})$, at that $N_1+N_2+...+N_M=N$.
+The subdomains $D_i, 1 \leq i \leq M$, are projections of initial search domain $D$ onto the subspaces corresponding to the variables $u_i, 1 \leq i \leq M$.
+
+\Russian 
+¬ проведенном исследовании данна€ схема примен€лась при M=2, т.е. использовалс€ только один уровень вложенности
+\begin{equation}
+\min_{y \in D}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}= \min_{u_1\in D_1}\min_{u_2\in D_2}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}.
+\end{equation}
+
+\section{Organization of Parallel Computing}\label{sec:4}
+
+\section{Numerical experiments}\label{sec:5}
+
+\begin{itemize}
+	\item one could control the size of the feasible domain $Q_{m+1}$ with respect to the whole domain $D$;
+	\item the global minimizer of the objective function is known a priori taking into account the constraints;
+	\item the global minimizer of the objective function without accounting for the constraints is out of the feasible domain $Q_{m+1}$  (with the purpose of simulating the behavior of the constraints and the objective function in the applied constrained optimization problems).
+\end{itemize}
+
+\begin{figure}[ht]
+\begin{minipage}[h]{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{vag_06.png} \\ (a)}
+\end{minipage}
+\hfill
+\begin{minipage}[h]{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{GKLS_04.png} \\ (b)}
+\end{minipage}
+\center{{\bf Figure 1.} The level lines of (a) hard and (b) simple subproblems }
+\end{figure}
+
+The numerical experiments were carried out using two classes of problems (\textit{Simple} ans \textit{Hard}) with $N=5$.
+The problem was considered to be solved if the algorithm generates a trial point $y^k$ in $\delta$-vicinity of the global minimizer, i.e.
+$\left\|y^k-y^\ast\right\|\leq\delta$ .
+The size of the vicinity was selected as $\delta=0.01\left\|b-a\right\|$, where $a$ and $b$ are the boundaries of the search domain $D$.
+The maximum allowable number of iterations was $K_{max} = 10^6$.
+
+\Russian
+ѕервый эксперимент проведем при решении серий задач Simple и Hard на одном узле, полностью задействовать оба доступных CPU (т.е.
+p=16 физических €дер).
+¬ таблице 1 приведено среднее врем€ (в секундах), потребовавшихс€ дл€ решени€ задач серии.
+
+\begin{table}[ht]
+\center{{\bf Table 1.} Average time for solving the problem on one node}
+\begin{center}
+\begin{tabular}{cc}
+\hline
+Problems & Time, sec. \\
+\hline
+\textit{Hard} & 52  \\
+\textit{Simple} & 51 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\Russian
+—ледующий эксперимент проведем, задействовав p узлов по три графических ускорител€ на каждом из них.
+“ак, при p=32 было задействовано 94 GPU ускорител€, на каждом из которых Ц 2688 CUDA-€дер.
+¬ таблице 2 приведено врем€ работы алгоритма, а в таблице 3 Ц ускорение по отношению к запуску на одном узле.
+
+\begin{table}[ht]
+\center{{\bf Table 2.} Average time for solving the problem on $p$ nodes}
+\begin{center}
+\begin{tabular}{cccc}
+\hline
+Problems & $p = 8$ & $p=16$ & $p=32$ \\
+\hline
+\textit{Hard} & 11.51 & 5.53 & 0.54 \\
+\textit{Simple} & 2.04 & 1.50 & 0.47\\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}[ht]
+\center{{\bf Table 3.} Speedup with respect to one node}
+\begin{center}
+\begin{tabular}{cccc}
+\hline
+Problems & $p = 8$ & $p=16$ & $p=32$ \\
+\hline
+\textit{Hard} & 5 & 9 & 96\\
+\textit{Simple} & 25 & 34& 109\\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+Computational experiments were carried out on a high-performance cluster of Lobachevsky State University of Nizhni Novgorod.
+The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM, and three NVIDIA Kepler K20’ GPUs (2688 CUDA cores, 6 Gb GDDR5).
+
+\section{Conclusion}
 

### 2017-11-06 2017/ParOptBook/Chapter2.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter4.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter5.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter6.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter7.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter8.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\dedic.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\editor.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\foreword.tex
D:\Common\Barkalov\Публикации\20

--- 2017-09-01 2017/ParOptBook/Chapter7.tex
@@
-%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% sample root file for your "contribution" to a contributed volume
-%
-% Use this file as a template for your own input.
-%
-%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
 %\documentclass[graybox]{svmult}
-%
-%% choose options for [] as required from the list
-%% in the Reference Guide
 %
@@
 %\usepackage{type1cm}        % activate if the above 3 fonts are
-                             % not available on your system
-%
+                            %% not available on your system
+%%
 %\usepackage{makeidx}         % allows index generation
 %\usepackage{graphicx}        % standard LaTeX graphics tool
-%                             % when including figure files
+                             %% when including figure files
 %\usepackage{multicol}        % used for the two-column index
-%\usepackage[bottom]{footmisc}% places footnotes at page bottom
+%\usepackage[bottom]{footmisc}
+%\usepackage{amsmath}
 %
-%% see the list of further useful packages
-%% in the Reference Guide
+%
 %
 %\makeindex             % used for the subject index
-%                       % please use the style svind.ist with
-%                       % your makeindex program
+                       %% please use the style svind.ist with
+                       %% your makeindex program
 %
@@
 %\begin{document}
+%
+\title{Information and characteristical global search algorithms}
+%\titlerunning{Global optimization via dimensionality reduction}
 
-\title{Software Systems for Global Optimization }
-% Use \titlerunning{Short Title} for an abbreviated version of
-% your contribution title if the original one is too long
-\author{Name of First Author and Name of Second Author}
-% Use \authorrunning{Short Title} for an abbreviated version of
-% your contribution title if the original one is too long
-\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
-\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
-%
-% Use the package "url.sty" to avoid
-% problems with special characters
-% used in your e-mail or web address
+\author{Vladimir A.Grishagin, Dmitri E.Kvasov, Marat A.Mukhametzhanov and Roman G.Strongin}
+\authorrunning{V.Grishagin, D.Kvasov, M.Mukhametzhanov, R.Strongin} 
+\institute{Vladimir A.Grishagin \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia   \email{vagris@unn.ru}
+\and Dmitri E.Kvasov \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia 
+\at DIMES, University of Calabria, Via P. Bucci, Cubo 42-C, 87036 Rende (CS), Italy \email{yaro@si.deis.unical.it}
+\and Marat A.Mukhametzhanov \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia 
+\at DIMES, University of Calabria, Via P. Bucci, Cubo 42-C, 87036 Rende (CS), Italy \email{yaro@si.deis.unical.it}
+\and Roman G.Strongin \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia   \email{vagris@unn.ru} }
+
 %
@@
 
-\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\abstract{This chapter describes the main principles of two approaches to development and investigation of global optimization algorithms.
+The first approach called statistical one considers a probabilistic model of the optimization problem in which the objective function is a realization of some stochastic process with predefined properties.
+In this model the optimization algorithm is built as an optimal stochastic procedure implementing a Bayesian strategy.
+A brief description of the general technique of deriving the methods within the framework of this approach is given.
+Some simple algorithmic schemes as a result of applying the techniques are presented.
+The second approach is called characteristical and focused on a general structure of decision rules for optimization algorithms.
+}
 
-\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+\section{Information approach to the global search}
+\label{sec:2_1}
+The information approach proposed by Prof.
+Strongin \cite{2_StrMonRus, 2_Str1989, 2_StrSergMon2000} provides a theoretical basis for the development of the global extremum search algorithms.
+At first, we will give a brief description of the general technique of deriving the methods within the framework of this approach.
+Further, some simple algorithmic schemes as a result of applying the technique will be presented.
 
-\section{Section Heading}
-Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.
+Consider the optimization problem 
+\begin{equation}
+\label{eq:2_-1}
+\varphi(y)\rightarrow\min, y \in Q\subseteq R^N.
+\end{equation}
+i.e., the problem of finding the lowest (minimal) value of the objective function $\varphi(y)$ and the coordinate (global minimizer) of this value in the admissible domain $Q$ of the N-dimensional Euclidean space $R^N$.
 
+According to the computational model formed within the framework of this approach it is supposed for the problem (\ref{eq:2_-1}) that a probability distribution $F$ is defined in the class of the measurable subsets of the set $\Phi$ of all functions $\varphi$ defined in the domain $Q$.
+The acceptance of this model means that the minimized function $\varphi$ is considered as a realization of some stochastic process, i.e., as the function of two arguments $\varphi=\varphi(y,\sigma)$, where $y\in Q$ and $\sigma$ is a random value, the properties of which are determined by the distribution $F$.
+
+Any additional information on the function $\varphi$ (known a priori as well as a posteriori) can be estimated in the probabilistic sense as a probability $F(\Phi\acute{})$  of the function $\varphi$ to belong to the set  $\Phi\acute{}$ of the functions possessing the properties corresponding to the given information.
+Particularly, the distribution $F$ is assumed a priori can be transformed to a posteriori distribution   with respect to the search information obtained in the course of optimization.
+This distribution allows estimating the position of extremum, predicting the result of the next trial, and designing the optimization algorithms based on the statistical decision rules.
+The detailed description of this methodology is given in the monographs ~\cite{2_StrMonRus} and \cite{2_StrSergMon2000}.
+Here we will demonstrate briefly an application of this approach to the one-dimensional case.
+
+Let in the problem (\ref{eq:2_-1})  the dimension $N=1$ and the admissible domain Q be a closed interval $[a,b]$ of the real axis.
+In order to emphasize that the problem under study is  one-dimensional  the variable $x$ will be used instead of the objective function argument $y$.
+Further, consider given problem
+\begin{equation}
+\label{eq:2_0}
+\varphi(x)\rightarrow\min, x \in Q.
+\end{equation}
+in a discrete statement.
+
+Let us define a finite uniform $\epsilon$-grid in the interval $[a,b]$ 
+\begin{equation}
+\label{eq:2_1}
+x_0=a<x_1=a+\epsilon<\ldots<x_i=a+i\epsilon<\ldots x_n=b
+\end{equation}
+and introduce the set of grid node indices 
+\begin{equation}
+\label{eq:2_2}
+U=\{0,1,\ldots n\}.
+\end{equation}
+Assume that in the problem (\ref{eq:2_0}) the set $Q=\{x_i,i\in U\}$ .
+Then it is necessary to determine the node $x_\nu$ , in which $\varphi(x)$ defined over the domain $Q$ reaches the global minimum, i.e.,
+\begin{equation}
+\label{eq:2_3}
+\varphi(x_\nu)\leq \varphi(x_i),i\in U.
+\end{equation}
+In the present model any function $\varphi(x)$  is defined by the values $\varphi_i=\varphi(x_i),0\leq i\leq n$  completely, i.e., it can be represented as a vector
+\begin{equation}
+\label{eq:2_4}
+\varphi=(\varphi_0,\ldots \varphi_n)\in R^{n+1}
+\end{equation}
+in $(n+1)$-dimensional Euclidean space $R^{n+1}$.
+Then, the a priori assumptions of the minimized function  $\varphi\in R^{n+1}$ could be described by defining the density $p(\varphi$) of a probabilistic distribution $F$ over the space $R^{n+1}$.
+
+If the density $p(\varphi)$ is  positive and continuous then the minimum of the optimized function is achieved at a single point with the probability equal to unity.
+Therefore, in order to estimate this minimum, it is sufficient  to compute the probabilities $\eta(\alpha)$ of minimum location at the points $\alpha$ coinciding with the grid nodes.
+Moreover, after executing the trials at some grid nodes and obtaining the search information $\omega_k$, one can deduce some conditional probabilities $\eta(\alpha/\omega_k)$  providing \textit{a posteriori} the current estimates of the extremum.
+
+Unfortunately, the computation of these probabilities requires an integration over the complicated multidimensional domains.
+To simplify the model, an additional random value  $\xi(\alpha)$ (called \textit{the nature state}) is introduced  which is connected with the given  probability density $p(\varphi)$  by the decomposition
+\begin{equation}
+\label{eq:2_5}
+p(\varphi)=\sum_{\alpha\in U}p(\varphi/\alpha)\xi(\alpha).
+\end{equation}
+The introduction of this decomposition allows establishing the nearness condition for the a posteriori probabilities $\eta(\alpha/\omega_k)$  and $\xi(\alpha/\omega_k)$  and, therefore, passing to the computation of the conditional probabilities of the nature state, which is simpler essentially.
+
+As a particular example of an application of the information approach, let us consider a probabilistic description, where the a priori known probability $p(\varphi)$ is represented by a decomposition $p(\varphi/\alpha),0\leq \alpha\leq n,$ where
+\begin{equation}
+\label{eq:2_6}
+p(\varphi/\alpha)=\prod_{i=0}^n {\frac{1}{\sigma_i\sqrt{2\pi}}\exp\left\{-\sum_{i=0}^n {\frac{(\Delta\varphi_i-m_i(\alpha))^2}{2\sigma_i^2}}\right\}}.
+\end{equation}
+Here
+\begin{displaymath}
+\Delta\varphi_i=\varphi_i-\varphi_{i-1},1\leq i\leq n,\; \Delta\varphi_0=\varphi_0,
+\end{displaymath}
+\begin{displaymath}
+\sigma_i=cm,\;1\leq i\leq n,\;c>0,\;m>0,
+\end{displaymath}
+\begin{displaymath}
+m_i(\alpha) =
+  \begin{cases}
+    -m, & i\leq\alpha, \\
+    \ \ \  m, & i>\alpha,
+  \end{cases} \ 1\leq i\leq n.
+\end{displaymath}
+
+According to the present description, for any fixed $\alpha$  the first differences $\Delta\varphi_i$  and the function value $\varphi_0$  at the point $x_0$ are the samples of the mutually independent Gaussian random variables with the standard deviations $\sigma_i=cm,1\leq i \leq n,\;\sigma_0$   and mathematical expectations $m_i(\alpha),1\leq i\leq n,\;m_0$.
+
+These assumptions are some probabilistic analog of Lipschitz condition, which bounds the first differences of the minimized function and provides its uniform continuity.
+Indeed, according to the description the variances of the differences $\Delta\varphi_i$  of the function $\varphi$ are bounded.
+Besides, these differences are the realizations of the independent normal values that in the case $mn=const$  provides a probabilistic property analogous to the uniform continuity.
+
+In addition,  for any given $\alpha$ the a priori mathematical expectations
+\begin{displaymath}
+\mu_\alpha(i) =m_0-m\times
+  \begin{cases}
+    i, & i\leq\alpha, \\
+    2\alpha-i, & i>\alpha,
+  \end{cases} 
+\end{displaymath}
+for the values $\varphi_i$  of the function $\varphi$  has a single local minimum, i.e., the functions close to the constants are unlikely.
+
+In the model considered an exclusion of the parameters $m_0$  and $\sigma_0$  determining the mathematical expectation and the standard deviation of $\varphi_0$  is possible if the initial trial is executed at the node with the index $i_0=0$ .
+In this case some asymmetry arises because of the function value at the left boundary is defined initially.
+In order to correct this asymmetry, the next trial should be executed at the right end of the grid, i.e., at the point with the index $i_1=n$ .
+
+Let us now assume that the trials executed already have been renumbered by subscripts in increasing order of the grid nodes 
+\begin{equation}
+\label{eq:2_7}
+a=i_0<i_1<\ldots<i_{k-1}<i_k=b,
+\end{equation}
+and that the information
+\begin{equation}
+\label{eq:2_8}
+\omega=\omega_k=\{(i_s,z_s),0\leq s\leq k\},
+\end{equation}
+where $z_s=\varphi(i_s),0\leq s\leq k,$  has been obtained.
+
+Then the nature state probabilities determined \textit{a posteriori} can be expressed as 
+\begin{displaymath}
+\xi(\alpha/\omega_k)=B\times h(\omega_k,\alpha)\xi(\alpha),
+\end{displaymath}
+where $B$ is a constant and for any $\alpha$ from the interval  $i_{s-1}\leq\alpha\leq i_s,1\leq s\leq k,$
+\begin{displaymath}
+h(\omega_k,\alpha)=\exp\left\{-\frac{(\alpha-\alpha_s^*)^2}{2\rho_s^2}\right\}\exp\left\{\frac{R(s)}{2mc^2}\right\},
+\end{displaymath}
+\begin{equation}
+\label{eq:2_9}
+\alpha_s^*=\frac{i_s-i_{s-1}}{2}-\frac{z_s-z_{s-1}}{2m},
+\end{equation}
+\begin{equation}
+\label{eq:2_10}
+R(s)=m\Delta_s+\frac{(z_s-z_{s-1})^2}{m\Delta_s}-2(z_s+z_{s-1}),
+\end{equation}
+\begin{displaymath}
+\rho_s^2=\left(\frac{c}{2}\right)\Delta_s,\ \Delta_s=i_s-i_{s-1}.
+\end{displaymath}
+Let us assume that the distribution  $\xi(\alpha)$ is a uniform one ($\xi(\alpha)=1/(n+1)$), i.e., all initial states of nature are of equal probability, and the parameter $m$ satisfies the condition
+\begin{equation}
+\label{eq:2_11}
+m>\max_{1\leq s\leq k}\left|\frac{z_s-z_{s-1}}{\Delta_s}\right|.
+\end{equation}
+Then, if $c$ is small enough, the most probable nature state is achieved at the point
+\begin{equation}
+\label{eq:2_12}
+\alpha^*=E(\alpha_t^*+0.5),
+\end{equation}
+of the interval $(i_{t-1},i_t)$ , for which the value $R(t)$  is maximal among all  the values $R(s)$  from (\ref{eq:2_10}).
+Here $E$ is the function of the integer part of its argument, i.e., $\alpha^*$  is the closest integer number to $\alpha_t^*$  from (\ref{eq:2_9}).
+Because of nearness of the a posteriori probabilities $\eta(\alpha/\omega_k)$  and  $\xi(\alpha/\omega_k)$ the point (\ref{eq:2_12}) is the most probable point of the global minimum.
+
+Note an important aspect connected with the parameter $m$ and the condition (\ref{eq:2_11}).
+In order to satisfy it, we will use the following adaptive scheme to estimate the parameter $m$ according to the results of the trials (\ref{eq:2_8}):
+\begin{equation}
+\label{eq:2_13}
+m =\begin{cases}
+    rM, & M>0, \\
+    1, & M=0,
+  \end{cases} 
+\end{equation}
+where
+\begin{equation}
+\label{eq:2_14}
+M=\max_{1\leq s\leq k}\left|\frac{z_s-z_{s-1}}{i_s-i_{s-1}},\right| 
+\end{equation}
+
+The coefficient $r$ is considered to be a parameter of the method and to satisfy the inequality $r>1$.
+
+Thus, in order to continue the minimum search after execution of the trials at the nodes (\ref{eq:2_7}), it seems naturally to select a new node with the maximum probability to find the sought extremum, i.e., the point (\ref{eq:2_12}).
+But in order to find this point it is enough to compute the values   from (\ref{eq:2_10}) called characeristics of the corresponding intervals   and to select the greatest of these ones.
+As a result, the complex problem of direct computation of the most probable minimum position via the multidimensional integration is reduced to a simple problem of the computation of characteristics (\ref{eq:2_10}).
+
+\section{The core global search algorithm}
+\label{sec:2_2}
+Let us consider the one-dimensional problem of the minimization of a function over an interval:
+\begin{equation}
+\label{eq:2_15}
+\varphi(x)\rightarrow \min,x\in Q=[a,b]. 
+\end{equation}
+
+Let us expand the discrete algorithm described in the above section onto the continuous case of the problem (\ref{eq:2_15}) by increasing the number of nodes infinitely ($n\rightarrow \infty$) so that finally the finite interval $Q=[a,b]$  is obtained.
+In this situation, the discrete algorithm can be transformed into its limiting continuous variant in the following way.
+
+In the relations (\ref{eq:2_10}), (\ref{eq:2_14}) the indices  from (\ref{eq:2_7}) are substituted by the corresponding grid nodes (\ref{eq:2_1}).
+
+\begin{equation}
+\label{eq:2_16}
+x_{i_s}=a+i_s((b-a)/n), 0\leq s\leq k. 
+\end{equation}
+This substitution does not change the values of the characteristics $R(s)$ in (\ref{eq:2_10}).
+
+Let us hereafter denote the node $x_{i_s}$ as $x_s$  for simplification. The substitution of (\ref{eq:2_16}) in (\ref{eq:2_9}) leads to the expression for the point of a new  $(k+1)$-th trial
+\begin{equation}
+\label{eq:2_17}
+x^{k+1}=a+\alpha_t^*((b-a)/n)=\frac{1}{2}(x_t-x_{t-1})-\frac{z_t-z_{t-1}}{2\tilde{m}},
+\end{equation}
+where $\tilde{m}$ is connected with $m$ from (\ref{eq:2_13}) as $\tilde{m}=m(\frac{b-a}{n})$.
+
+Note that the point $x_{\alpha^*}=a+\alpha^*(\frac{b-a}{n})$  where $\alpha^*$  from (\ref{eq:2_12}) is the node of the grid (\ref{eq:2_1}) closest to the point (\ref{eq:2_17}).
+This means that if $n\rightarrow\infty$ the point $x_{\alpha^*}$  tends to the point $x^{k+1}$.
+
+Hereinafter the «continuous» algorithm described above will be called as \textit{the core algorithm of the global search (AGS)}.
+
+\subsection {Computational scheme of the core algorithm}
+\label {subsec:2_2_1}
+Let us give a detailed description of the AGS computational scheme applied to solving the problem (\ref{eq:2_15}) and consider the set 
+\begin{equation}
+\label{eq:2_18}
+\omega=\omega_k=\{(x^i,z^i),1\leq i\leq k\}
+\end{equation}
+as the search information after $k$ trials an points $x^i$ with results $z^i=\varphi(x^i),\:1\leq i\leq k$.
+According to the algorithm, the first two trials are executed at the ends of the interval  $[a,b]$, i.e., $x^1=a,\;x^2=b$.
+The function values $z^1=\varphi(x^1),\;z^2=\varphi(x^2)$  are computed and the number $k$ of the executed trials is set to 2.
+
+Assume that $k\geq 2$  trials have been executed and that the information (\ref{eq:2_18}) has been obtained.
+In order to select the point of a new trial $x^{k+1}$  it is necessary to perform the following steps.
+\begin{description}[\textbf{Step 1}]
+\item[\textbf{Step 1}]{Renumber by subscripts (beginning from zero) the points $x^i,\:1\leq i\leq k$,  from (\ref{eq:2_18}) in increasing order, i.e.,
+\begin{displaymath}
+a=x_0<x_1<\ldots <x_{k-1}=b.
+\end{displaymath} }
+\item[\textbf{Step 2}]{Juxtapose to the points $x_i,\:1\leq i\leq k-1$,  the function values $z_i=\varphi(x_i),\:1\leq i\leq k-1$,  corresponding to trial results $z^i$  from (\ref{eq:2_18}) and renumbered by subscripts.}
+\item[\textbf{Step 3}]{Compute the maximal absolute value of the first divided differences 
+\begin{displaymath}
+M=\max_{1\leq i\leq k-1}\left|\frac{z_i-z_{i-1}}{x_i-x_{i-1}}\right|
+\end{displaymath}
+and accept
+\begin{equation}
+\label{eq:2_19}
+m =\begin{cases}
+    rM, & M>0, \\
+    1, & M=0,
+  \end{cases} 
+\end{equation}
+where $r>1$ is a predefined parameter of the method. } %end of Step 3
+\item[\textbf{Step 4}]{For each interval $(x_{i-1},x_i),1\leq i\leq k-1$,  calculate the characteristic
+\begin{displaymath}
+R(i)=m(x_i-x_{i-1})+\frac{(z_i-z_{i-1})^2}{m(x_i-x_{i-1})}-2(z_i+z_{i-1}).
+\end{displaymath} } %end of Step 4
+\item[\textbf{Step 5}]{Find the interval $(x_{t-1},x_t)$ , which the maximum characteristic
+\begin{equation}
+\label{eq:2_20}
+R(t)=\max_{1\leq i\leq {k-1}}R(i).
+\end{equation}  } %end of Step 5
+\item[\textbf{Step 6}]{Execute the new trial at the point 
+\begin{displaymath}
+x^{k+1}=\frac{1}{2}(x_{t-1}+x_t) - \frac{z_t-z_{t-1}}{2m}.
+\end{displaymath}
+compute the value $z^{k+1}=\varphi(x^{k+1})$ , and increment the search step counter: $k=k+1$  .
+} %end of Step 6
+\end{description}
+
+Following the general scheme of optimization algorithms ~\cite{2_GriSergChap1}  the operations of Steps 1-6 describe the decision rule of AGS.
+
+The termination criterion is given as
+\begin{equation}
+\label{eq:2_21}
+H_k(\Phi,\omega_k) =
+  \begin{cases}
+    0, & x_t-x_{t-1}\leq\epsilon, \\
+    1, & x_t-x_{t-1}>\epsilon,
+  \end{cases}
+\end{equation}
+where $\epsilon>0$  is a predefined search accuracy.
+
+Finally, the pair
+\begin{equation}
+\label{eq:2_21_1}
+e^k=(\varphi_k^*,x_k^*)
+\end{equation}
+where  $\varphi_k^*$ is the lowest of the computed function values, i.e.,
+\begin{equation}
+\label{eq:2_21_2}
+\varphi_k^*=\min_{1\leq i\leq k}\varphi(x^i)=\min_{0\leq i\leq {k-1}}\varphi(x_i),
+\end{equation}
+and
+\begin{displaymath}
+x_k^*=\arg \min_{1\leq i\leq k}\varphi(x^i),
+\end{displaymath}
+(the coordinate of $\varphi_k^*$ ), is taken as the estimate of the minimum.
+
+The most important theoretical properties of AGS concerning convergence and efficiency of the method are given in the next subsection. 
+\subsection {Conditions of convergence and efficiency estimates}
+\label{subsec:2_2_2}
+This subsection contains a brief description of the main theoretical results regarding the properties of the core information algorithm.
+The strict substantiation of these properties has been given in the monographs ~\cite{2_StrMonRus, 2_StrSergMon2000} and can be derived as a consequence of the general characteristical theory of global optimization methods ~\cite{2_GrishaginSergeyevStrongin}.
+
+Let us consider an infinite sequence of trials $\{x^k\}=\{x^1,x^2,\ldots ,x^k,\ldots\}$  generated by AGS in the absence of the termination criterion.
+Since $\{x^k\}\subseteq [a,b]$  this sequence contains necessarily at least one converging subsequence.
+Assume that $x^+$  is the accumulation point (the limit of some converging subsequence) of the sequence $\{x^k\}$ .
+Then
+\begin{description}[i)]
+\item [i)]{if $a<x^+<b$ ($x^+$ is an internal point of the closed interval $[a,b]$)  the trial sequence contains two subsequences, one of which converges to $x^+$  from the left and the other – from the right;}
+\item [ii)]{for any $k\geq 1$  the inequality $\varphi(x^k)\geq \varphi(x^+)$ is true;}
+\item [iii)]{if there exists another limit point $x^{++}$  of the sequence $\{x^k\}$, then $\varphi(x^{++})=\varphi(x^+)$;}
+\item [iv)]{if the function $\varphi(x)$ has a finite number of local extrema within the interval $[a,b]$ then the point $x^+$ is a local minimum point;}
+\item [v)]{if the function $\varphi(x)$ satisfies the Lipschitz condition in the interval $[a,b]$ with a constant $L>0$, i.e.,
+\begin{equation}
+\label{eq:2_22}
+\left|\varphi(x')-\varphi(x'')\right|\leq L\left|x'-x''\right|
+\end{equation} 
+and, beginning from some search step, the inequality 
+\begin{equation}
+\label{eq:2_23}
+m>2L
+\end{equation}
+holds then any point of the global minimum of the function $\varphi(x)$ in the interval $[a,b]$ is a limit (accumulation) point of the AGS trial sequence.} 
+\end{description}
+
+As a consequence of  assertions (iii) and (v), under conditions (\ref{eq:2_22}) and (\ref{eq:2_23}) AGS finds all the global minimum points of a Lipschitzian objective function and converges only to such points.
+This means that the method generates an essentially non-uniform trial grid in the search domain.
+In order to estimate this non-uniformity reflecting the efficiency of the trial placement by the algorithm, let us introduce the concept of the trial density as a measure of this efficiency.
+\begin{definition}
+\label{def:2_1}
+\textit{The density of the sequence} $\{x^k\}$   in the subinterval $[\alpha,\beta]\subset [a,b]$  is the quantity
+\begin{displaymath}
+P_{\alpha\beta}=\frac{K_{\alpha\beta}}{\beta-\alpha},
+\end{displaymath}
+where $K_{\alpha\beta}$ is the number of points of the trial sequence $\{x^k\}$  falling into the interval $[\alpha,\beta]$.
+\end{definition}
+
+Assume that a Lipschitzian (\ref{eq:2_22}) function $\varphi(x)$ satisfies the following inequality over an interval $[\alpha,\beta]$. 
+\begin{equation}
+\label{eq:2_24}
+\varphi(x)\geq\varphi(x^*)+\Delta,\;x\in [\alpha,\beta],
+\end{equation}
+where  $\Delta>0$ and $x^*$  is the global minimum point of $\varphi(x)$ in the interval $[a, b]$.
+If $\beta-\alpha>2\Delta/L$  and the condition $m\geq 2L$  is satisfied, then for AGS ~\cite{2_StrMonRus, 2_StrSergMon2000}
+\begin{equation}
+\label{eq:2_25}
+P_{\alpha\beta}\leq\frac{3m}{2\Delta}.
+\end{equation}
+
+Note that the subinterval $[\alpha,\beta]$ satisfying (\ref{eq:2_24}) can contain no global minimum.
+It means that in this subinterval a finite number of trials only can be located as a consequence of the statement (iii).
+The inequality (\ref{eq:2_25}) gives a quantitative estimate of this number.
+
+Moreover, the inequality (\ref{eq:2_25}) allows comparing the trial density for AGS and for the uniform scanning method.
+If the global minimum value $\varphi(x^*)$   in the problem (\ref{eq:2_15}) with the Lipschitzian function (\ref{eq:2_22}) can be found with a precision $\gamma>0$  (with respect to the function value) by the  method of scanning  the uniform grid nodes then the step of this grid should not exceed the value $2\gamma/L$  in order to achieve the precision $\gamma$.
+In this case, the trial density $\tilde{P_{\alpha\beta}^*}$ of the scanning scheme over the interval $[\alpha,\beta]$ satisfies the inequality
+\begin{equation}
+\label{eq:2_26}
+P_{\alpha\beta}^*\geq \frac{L}{2\gamma}.
+\end{equation}
+
+If the Lipschitz constant $L$ is known, let us assume for AGS that $m=2L$ . Then from (\ref{eq:2_25}) and (\ref{eq:2_26})
+\begin{equation}
+\label{eq:2_27}
+P_{\alpha\beta}\leq \left(\frac{6\gamma}{\Delta}\right)P_{\alpha\beta}^*.
+\end{equation}
+It follows from (\ref{eq:2_27}) that the higher the required precision $\gamma$ of the problem solution, the higher the degree of the efficiency of AGS as compared to the scanning method.
+
+Let us give also the theoretical comparison of AGS with the broken lines method proposed by Piyavskij ~\cite{2_Piyavskij} for the minimization of the Lipschitzian functions (\ref{eq:2_22}).
+This method like AGS uses some estimate $m$ for the Lipschitz constant  $L$ as their parameter.
+Let us assume that this estimate is the same for both the methods and that the objective function is linear over the interval $[\alpha,\beta]$.
+Then the inequality
+\begin{displaymath}
+P_{\alpha\beta}\leq 2P_{\alpha\beta}^+,
+\end{displaymath}
+takes place ~\cite{2_GrishaginDynSys} where $P_{\alpha\beta}^+$  is the density of Piyavskij method over the interval $[\alpha,\beta]$.
+\subsection {Operational characteristics}
+\label{subsec:2_2_3}
+The complexity of the investigated problem and of the applied methods allows obtaining the theoretical estimates of efficiency and performing the theoretical comparison of the algorithms in rather rare cases only.
+In this situation, the numerical experiments become an important element of analyzing the algorithms.
+To compare various algorithms to each other, the authors have proposed a technique based on building the operational characteristics of the methods ~\cite{2_GrishaginOperChar, 2_StrSergMon2000}.
+Let us give a brief description of this approach to the comparison of the algorithms’ efficiencies.
+
+The efficiency of any optimization algorithm is connected inseparably with two criteria, namely,
+
+\begin{enumerate}
+\item {the computational cost of the search to the moment of termination;}
+\item {the precision of the problem solution determined by the extremum estimate $e^k$  from (\ref{eq:2_21_1})}. 
+\end{enumerate}
+
+Assume that the search cost is measured in the number of trials $K$ executed to the moment of termination and the accuracy is determined by the value $P=\varphi_K^*-\varphi^*$   where $\varphi_K^*$  from (\ref{eq:2_21_2}) for $k=K$ is the lowest value of the function computed after $K$ steps and  $\varphi^*$ is the minimum value of the objective function.
+Note that the criteria $K$ and $P$ are the contradictory ones, namely, the better the accuracy, the larger the computational costs and vice versa, the less the number of trials, the worse the precision of the problem solution.
+In other words, introducing these characteristics, we formulate a bi-criterial optimization problem with contradictory criteria.
+
+Having completed the numerical solving of a particular problem with given set of the method parameters, one can obtain the particular values of the criteria, which are convenient to present as a point on a coordinate plane with the axes $K$ and  $P$.
+A set of such points corresponding to different combinations of the algorithm parameters is called \textit{the operational characteristic} of the method.
+
+The operational characteristics obtained via the procedure described above reflects the results of the optimization of one function only.
+The results connected with the functions belonging to some function class are much more informative.
+Let us expand the concept of the operational characteristic onto this case.
+
+Let us consider a class $\Psi$ of the functions $\psi$ and assume that a probability distribution $F(\Psi)$  is defined over this class.
+Let us select randomly $n$ functions $\psi_i\in\Psi,\:1\leq i\leq n,$  according to this distribution and perform their optimization with given parameters of the method.
+After the experiment for the $i$-th function we have the number $K_i$  of trials executed and the accuracy $P_i$ .
+Let us take the averaged number of trials
+\begin{displaymath}
+K=\sum_{i=1}^n{K_i}/n
+\end{displaymath}
+and the probability $P$ of finding the global minimum to the moment of the experiment termination with given precision $\delta>0$, i.e.,
+\begin{displaymath}
+P=\sum_{i=1}^n{p_i}/n
+\end{displaymath}
+where
+\begin{displaymath}
+p_i =
+  \begin{cases}
+    1, & P_i<\delta, \\
+    0, & P_i<\delta,
+  \end{cases}
+\end{displaymath}
+as the criteria of the experiment efficiency. The probability $P$ will be called hereafter as \textit{the reliability} of the method.
+
+Having performed the optimization of a sample of functions with several method parameters, we obtain an approximate operational characteristics of the algorithm  for the class $\Psi$.
+
+Plotting the operational characteristics on the plane $(K,P)$   enables to compare visually the efficiency of different optimization algorithms.
+Namely, if the operational characteristic of one method for given $K$ is located above the one of another method, then the first method is more efficient since it provides a higher reliability for the same computational expenditures.
+Analogously, if the operational characteristic of one method for given value $P$ is positioned to the left from the operational characteristic of another one then the first method is better because it has spent less number of trials $K$ for solving the same number of problems.
+
+Note that, in general, the operational characteristic of a method is a set of points in the coordinate plane $(K, P)$.
+In this set a single value $K$ may correspond to several values of $P$ obtained for different method parameters.
+In this situation, the operational characteristics of different methods can overlap with each other, and in this overlap, one method can be better for particular method parameters while another one can be better with other parameter selection.
+How one can compare the methods to each other generally in this case?
+Since the method parameters could be selected according to researcher’s decision, it is proposed to compare the algorithms in accordance with the best operational characteristic indicators achieved, i.e., to compare the upper envelopes of the operational characteristics of the methods.
+
+Let us formalize this approach.
+Assume that the set $C=\{K(\pi),P(\pi)\}$ is the operational characteristic of an optimization algorithm obtained for groups of parameters $\pi\in\Pi$, where  $\Pi$  is the set of all possible combinations of method parameters.
+Let us construct the sections $C(K)$ of the set $C$ for different fixed $K$ as
+\begin{displaymath}
+C(K)=\{P:(K,P)\in C\}.
+\end{displaymath}
+and select the maximal reliabilities for all nonempty sections $C(K)$ 
+\begin{equation}
+\label{eq:2_28}
+P^*(K)=\max \{P:P\in C(K)\}.
+\end{equation}
+
+The curve (\ref{eq:2_28}) will be hereafter called \textit{the best operational characteristic} of the optimization method.
+Now the comparison of the efficiencies of the methods can be reduced to the comparison of their best operational characteristics.
+
+The condition $P(\pi)=P^*(K)$  determines the best method parameters for given $K$ and in such a way one can construct the set  $\pi^*$ of the optimal parameters  corresponding to the curve (\ref{eq:2_28}) for the optimization algorithm investigated on the chosen test class.
+
+The concept of operational characteristics has been recently expanded to the class of non-deterministic optimization methods as a more general concept of operational zones ~\cite{2_SergKvasMukhOZ}.
+
+Examples of operational characteristics for several algorithms  on different classes of test functions can be found, for example, in ~\cite{2_GerGriIsr, 2_GerGriGer, 2_GrishaginOperChar, 2_GriIsrSergAMC, 2_StrMonRus, 2_StrSergMon2000}.
+
+\section{Characteristical computational scheme for global search algorithms}
+\label{sec:2_3}
+The convergence to the sought solution (to the global minimum of the objective function of the problem (\ref{eq:2_15}) in the case considered) is an important property of a numerical method.
+The character of convergence determines the efficiency of the optimization method essentially.
+The present section is dedicated to the considering from a general theoretical viewpoint the convergence properties for a wide class of the numerical methods of the global extremum search for the functions of a single argument.
+This class called  \textit{the class of characteristical algorithms} includes many well-known methods  developed in the frameworks of various approaches to the design of optimization algorithms.
+
+Since many algorithms of multidimensional global search are based on the reduction of a multidimensional problem to a one-dimensional subproblem (or to a family of the one-dimensional ones) ~\cite{2_CarrHowe, 2_GerGriGer, 2_Goertzel, 2_SergStrLeraMonogr, 2_ShiOlaf, 2_StrSergMon2000} this theory can be applied to the analysis of multidimensional optimization problems as well.
+
+Consider a one-dimensional problem (\ref{eq:2_15}) with the search domain $Q=[a,b]$  and  optimization methods for solving this problem generating an infinite trial sequence $\{x^k\}=x^1,\:x^2,\:\ldots\;x^k,\ldots$, i.e., the methods in which the termination criterion is absent.
+
+\begin{definition}
+\label{def:2_2}
+A method for solving the problem (\ref{eq:2_15}) is called \textit{characteristical algorithm}  if, starting from some search step $k_0\geq 1$ , the choice of the coordinate $x^{k+1}$  of the next trial $(k\geq k_0)$  consists in the execution of the following steps.
+\begin{description}[\textbf{Step 1.}]
+\item [\textbf{Step 1}]{Construct the set 
+\begin{equation}
+\label{eq:2_29}
+\Lambda_k=\{x_0,\:x_1,\:\ldots\;,x_\tau\}
+\end{equation}
+of a finite number $\tau+1=\tau(k)+1$  of points from the domain $Q=[a,b]$  assuming that $a\in\Lambda_k$, $b\in\Lambda_k$, all the coordinates of the preceding trials $x^i\leq\Lambda_k,\;1\leq i\leq k,$ , and the elements of the set $\Lambda_k$  are renumbered by subscripts in the increasing order of the coordinate, i.e.,
+\begin{equation}
+\label{eq:2_30}
+a=x_0<x_1<\:\ldots\:<x_{\tau-1}<x_\tau=b.
+\end{equation}
+}%end of Step 1
+\item [\textbf{Step 2}]{Juxtapose to each interval $(x_{i-1},\:x_i),\;1\leq i\leq \tau$, a number $R(i)$ called  \textit{characteristic}  of this interval.
+}%end of Step 3
+\item [\textbf{Step 3}]{Select the interval $(x_{t-1},\:x_t)$ corresponding to the maximal characteristic 
+\begin{equation}
+\label{eq:2_31}
+R(t)=\max\{R(i):1\leq i\leq \tau\}.
+\end{equation}
+}%end of Step 3
+\item [\textbf{Step 4}]{Execute the next trial at a point
+\begin{equation}
+\label{eq:2_32}
+x^{k+1}=d(t)\in (x_{t-1},\:x_t).
+\end{equation}
+}%end of Step 4
+\end{description}
+\end{definition}
+According to the definition, the characteristical scheme of the algorithm defines the structure of its decision rule (see the general scheme of optimization methods in  ~\cite{2_GriSergChap1}) via the sequence of the operations described at Steps 1-4.
+The following meaningful interpretation can be given for these operations.
+
+In order to execute the next trial, the search domain $[a,b]$ is divided into $\tau$ subintervals $(x_{i-1},\:x_i),\:1\leq i\leq \tau$, by the points of the set $\Lambda_k$ .
+Next, for each interval the possibility to find out the global minimum in this interval is estimated (the interval characteristic is a mesure of such the possibility) and the interval with the best (maximal) characteristic is taken.
+The point of the next trial is selected in this interval according to the rule $d(\bullet)$.
+
+Note that the set (\ref{eq:2_29}) along with the trial coordinates may include the points, in which the trials have not been executed (for example, in a number of the information algorithms from ~\cite{2_StrSergMon2000} the ends of the search interval are such points).
+Concerning numbering, the upper index of trial coordinates corresponds to the order of the trial execution in the search process and the lower index determines the position of the point in the ordered array (\ref{eq:2_30}).
+Thus, the coordinate of the $i$-th  trial $x^i$  in the set $\Lambda_k$  gets a lower index  $j$, i.e., $x^i=x_j$  and index $j$ can change from one step to another $(j=j(k))$ .
+
+The concept of the characteristical structure of optimization methods has been introduced for the first time by Grishagin in ~\cite{2_GrishaginCharAlg}.
+Later, it was generalized and expanded onto other classes of problems and types of algorithms ~\cite{2_GrishaginSergeyevStrongin, 2_Pinter, 2_SergDivBest}.
+
+As examples, some well-known global search algorithms can be described in the characteristical form.
+In the first four algorithms, two initial trials are executed at the points $x^1=a$  and $x^2=b$ , the characteristical rule takes its power starting from $k \geq 2$.
+Morover, the set $\Lambda_k (k\geq2)$    consists of the trial points only, i.e.,  $\Lambda_k=\{x^1,x^2,\:\ldots\:,x^k\}$ and, consequently, $\tau=k-1$ .
+Let us also use the designation $z_j=\varphi(x_j)$  for the objective function values at the points $x_j\in \Lambda_k$ .
+\paragraph{\textbf{Method of sequential scanning}}
+In this method, the length of an interval is its characteristic, i.e.,
+\begin{equation}
+\label{eq:2_33}
+R(i)=x_i-x_{i-1}.
+\end{equation}
+and the next trial point is selected in the middle of the longest interval
+\begin{displaymath}
+x^{k+1}=\frac{1}{2}(x_t+x_{t-1}).
+\end{displaymath}
+\paragraph{\textbf{\textbf{Kushner's method \cite{2_Kushner}}}}
+The characteristic of the method has the form
+\begin{equation}
+\label{eq:2_34}
+R(i)=-\frac{4(\varphi_k^*-\delta_k-z_i)(\varphi_k^*-\delta_k-z_{i-1})}{x_i-x_{i-1}}
+\end{equation}
+and the next trial is executed at the point
+\begin{equation}
+\label{eq:2_35}
+x^{k+1}=x_{t-1}+\frac{(x_t-x_{t-1})(\varphi_k^*-\delta_k-z_t)}{2(\varphi_k^*-\delta_k)-z_t-z_{t-1}}.
+\end{equation}
+where $\delta_k>0$  is a parameter of the method, in general, depending on the number of the search step $k$,  and $\varphi_k^*$ from (\ref{eq:2_21_2}) is the lowest computed value of the objective function.
+\paragraph{\textbf{Method of broken lines}}
+In this method, which has been proposed by Piyavskij  ~\cite{2_Piyavskij} and later by Shubert ~\cite{2_Shubert} for the optimization of the Lipschitzian function (\ref{eq:2_22}).
+The characteristic of this method is calculated as
+\begin{equation}
+\label{eq:2_36}
+R(i)=\frac{m(x_{i-1}-x_i)}{2}-\frac{z_{i-1}+z_i}{2},
+\end{equation}
+and the point of the next trial is selected according to the expression
+\begin{equation}
+\label{eq:2_37}
+x^{k+1}=\frac{x_{t-1}+x_t}{2}-\frac{z_t-z_{t-1}}{2m},
+\end{equation}
+where $m>0$  is a parameter of the method.
+\paragraph{\textbf{Core information algorithm of global search (AGS)}}
+This method has been proposed by Strongin ~\cite{2_StrMonRus, 2_StrSergMon2000} as Bayesian one-step optimal algorithm and uses the characteristic
+\begin{equation}
+\label{eq:2_38}
+R(i)=m(x_i-x_{i-1})+\frac{(z_i-z_{i-1})^2}{m(x_i-x_{i-1})}-2(z_i+z_{i-1})
+\end{equation}
+The next trial point is selected according to (\ref{eq:2_37}). 
+The value $m>0$  is computed according to the expression
+\begin{equation}
+\label{eq:2_39}
+m =\begin{cases}
+    rM, & M>0, \\
+    1, & M=0,
+  \end{cases} 
+\end{equation}
+where
+\begin{equation}
+\label{eq:2_40}
+M=\max_{1\leq i\leq \tau}\left|\frac{z_i-z_{i-1}}{x_i-x_{i-1}}\right|
+\end{equation}
+and $r>1$ is a parameter of the method.
+\paragraph{\textbf{Information algorithm of global search with  trials at the internal points only (AGSIP)}}
+This algorithm ~\cite{2_SergStrAGSIP} is applied to the optimization of the functions over an interval without computations of the function values at the boundary points of the interval.
+This is important in the case, when the interval boundaries cannot be expressed explicitly.
+
+According to AGSIP, one or several initial trials are executed at internal points of the interval $(a,b)$.
+Then, the characteristical scheme comes into effect.
+The set $\Lambda_k$  includes the trial points and the ends of the interval only, from where $\tau=k+1$ .
+As before, $z_j=\varphi(x_j)$, however, for $1\leq j\leq k$  only.
+
+The characteristic of the method is computed according to the expression
+\begin{equation}
+\label{eq:2_41}
+R(i) =\begin{cases}
+    m(x_i-x_{i-1})+\frac{r(z_i-z_{i-1})^2}{M(x_i-x_{i-1})}-2(z_i+z_{i-1}), & 1<i<\tau, \\
+    m(x_1-x_0)-4z_1, & i=1, \\
+		m(x_\tau-x_{\tau-1})-4z_\tau, & i=\tau,
+  \end{cases} 
+\end{equation}
+and the next trial point is computed as 
+\begin{equation}
+\label{eq:2_42}
+x^{k+1} =\frac{x_{t-1}+x_t}{2}-
+\begin{cases}
+    0, & t=1\ or\ t=\tau, \\
+    (z_t-z_{t-1})(2m), & 1<t<\tau,
+  \end{cases} 
+\end{equation}
+where the value $m$ is taken according to (\ref{eq:2_39}) with 
+\begin{displaymath}
+M=\max_{2\leq i\leq \tau-1}\left|\frac{z_i-z_{i-1}}{x_i-x_{i-1}}\right|.
+\end{displaymath}
+After the particular examples, let us consider an interesting theoretical result on the connection of the one-step optimality principle of the optimization algorithms with the property of  characteristical decomposition.
+\begin{theorem} 
+\label{theor:2_1}
+One-step optimal (minimax or Bayesian) algorithm is the characteristical one.
+\end{theorem}
+\begin{proof}
+According to the one-step optimality principle ~\cite{2_Mockus, 2_StrMonRus, 2_StrSergMon2000} a real function $W_k(x)$  characterizing the efficiency of the trial execution at the point $x$ at the step $k$ is introduced, and it is supposed that the less the value of the criterion $W_k$ , the higher the efficiency.
+If an algorithm selects the next trial point $x^{k+1},\:k\geq k_0\geq 1$, according to the condition
+\begin{equation}
+\label{eq:2_43}
+W_k(x^{k+1})=\max_{x\in [a,b]}W_k(x)
+\end{equation}
+then this method is called \textit{one-step optimal algorithm}.
+
+Let us rewrite the condition of the one-step optimality (\ref{eq:2_43}) in the form 
+\begin{displaymath}
+W_k(x^{k+1})=\max_{1\leq i\leq \tau}\max_{x\in [x_{i-1},x_i]}W_k(x).
+\end{displaymath}
+Hence, the value
+\begin{displaymath}
+R(i)=\max_{x\in [x_{i-1},x_i]}W_k(x)
+\end{displaymath}
+may be taken as the characteristic of the interval $(x_{i-1},\:x_i)$ , and
+\begin{displaymath}
+d(t)=\arg \max_{x\in [x_{i-1},x_i]}W_k(x).
+\end{displaymath}
+The theorem has been proved.
+\end{proof}
+
+All the information univariate algorithms belong to the characteristical class since either these algorithms are derived from the principle of one-step optimality (like AGS) or they are generalizations of the core one-step optimal method inheriting the characteristical structure of their decision rules (like AGSIP).
+It should be noticed, that the method of broken lines and the Kushner's method were constructed as one-step optimal algorithms as well.
+
+General theoretical results on the convergence of the characteristical algorithms are presented in ~\cite{2_GrishaginCharAlg, 2_GrishaginSergeyevStrongin}.
+These results determine conditions providing the following properties:
+\begin{description}[i)]
+\item [i)] {bilateral convergence to internal limit points of trial sequence that allows to substantiate the  termination criterion for characteristical algorithms;}
+\item [ii)] {everywhere dense convergence;}
+\item [iii)] {convergence to locally optimal points;}
+\item [iv)] {convergence to all global minimizers (sufficient conditions of global convergence to the solution defined by the Problem statement C from Chapter 1).}
+\end{description}
+
+The detailed proof of assertions i) – iv) can be found in ~\cite{2_GrishaginSergeyevStrongin} and in this monograph (see ~\cite{2_GriSergChap3}) as a particular case of the more general results obtained for the class of parallel characteristical algorithms.
+
+\section{Speed up of global search on the base of additional information}
+\label{sec:2_4}
+The information global search algorithms considered in the present chapter are aimed at the solving of the multiextremal optimization problems with Lipschitzian functions.
+Constructing the optimizing sequences is performed sequentially and uses the search information $\omega_k$ obtained during the process of computations.
+Moreover, in order to solve the optimization problems the precise values of Lipschitz constants for the model functions are not required a priori.
+These values are estimated in the course of global search on the base of the search information available (see, for example, (\ref{eq:2_19}).
+Nevertheless, another one important issue remains, which requires a special attention for the design of the efficient decision rules of the global search algorithms.
+This issue consists in the fact that in most cases the behavior of the functions in the multiextremal global optimization problems is a non-uniform one in different subdomains of the search domain.
+In some subdomains the model function values can vary rather fast (that would correspond to high values of Lipschitz constant in these subdomains).
+In other subdomains (for example, in  vicinities of the extrema points) the function values can change more slowly.
+As a result, the integral (or common) estimates of the characteristics of the optimized function for the whole search domain can not correspond in full to the function properties in different subdomains of the search domain.
+Two directions may be outlined among the possible approaches taking into account the nonuniformity of the optimized function behavior in different subdomains of the search domain, in the framework of which the most valuable results have been obtained, namely:
+\begin{itemize}
+\item 
+Construction of separate estimates of Lipschitz constants for the optimized function in different subdomains of the search domain;
+\item 
+The use of the additional information on the optimized function behavior.
+In most cases, the values of derivatives at the points of the executed trials of the global search are considered as such additional information.
+ \end{itemize}
+Below we consider these approaches  in more details for the one-dimensional optimization problems.
+
+% 
+\begin{thebibliography}{99.}
+
+\bibitem{2_CarrHowe}	Carr, C.R., Howe, C.W.: Quantitative Decision Procedures in Management and Economic: Deterministic Theory and Applications. McGraw-Hill, New York (1964)
+\bibitem{2_GerGriIsr} Gergel, V.,  Grishagin, V., Israfilov, R.: Local tuning in nested scheme of global optimization, Procedia Computer Science \textbf{51}, 865--874 (2015) 
+\bibitem{2_GerGriGer}	Gergel, V., Grishagin, V., Gergel, A.: Adaptive nested optimization scheme for multidimensional global search, J. Glob. Opt. \textbf{66}, 35--51 (2016)
+\bibitem{2_Goertzel}	Goertzel, B.: Global Optimization with Space-Filling Curves. Appl. Math. Lett. 12, 133--135 (1999)
+\bibitem{2_GrishaginOperChar}  Grishagin,V.A.: Operation characteristics of some global search algorithms, Problems of Statistical Optimization \textbf{7},  198--206 (1978). In Russian
+\bibitem{2_GrishaginCharAlg} Grishagin, V.A.: On convergence conditions for a class of global search algorithms.
+In: Proc.
+of the 3rd All–Union Seminar on Numerical Methods of Nonlinear Programming,  82–84.
+Kharkov (1979).
+In Russian
+\bibitem{2_GrishaginDynSys} Grishagin, V.A.: Comparative investigation of search dynamics for two global optimization algorithms. System Dynamics \textbf{18}(3), 76--89 (1979). In Russian
+\bibitem{2_GriIsrSergAMC}	Grishagin, V.,  Israfilov, R., Sergeyev, Y.: Convergence conditions and numerical comparison of global optimization methods based on dimensionality reduction schemes.
+Appl.
+Math.
+Comput.
+\textbf{318}, 270--280 (2018)
+\bibitem{2_GrishaginSergeyevStrongin} Grishagin, V.A., Sergeyev, Y.D., Strongin, R.G.: Parallel characteristical algorithms for solving problems of global optimization.
+J.
+Global Optim.
+\textbf{10}(2), 185--206 (1997)
+\bibitem{2_GriSergChap1} Grishagin, V.A., Sergeyev, Y.D.: Global optimization: from one-dimensional to multidimensional problems via the dimensionality reduction. This volume.
+\bibitem{2_GriSergChap3} Grishagin, V.A., Sergeyev, Y.D.: Parallel algorithms for one-dimensional multiextremal optimization. This volume.
+\bibitem{2_Kushner}	Kushner, H.J.: A new method of locating the maximum point of an arbitrary multipeak curve in the presence of noise.
+Transactions of ASME, Ser.
+D.
+Journal of Basic Engineering \textbf{86}, 97--106 (1964)
+\bibitem{2_Mockus}	Mockus, J.: Bayesian Approach to Global Optimization. Kluwer Academic Publishers, Dordrecht (1988)
+\bibitem{2_Pinter}	Pint{\'e}r, J.D.: Global Optimization in Action.: Kluwer Academic Publishers, Dordrecht (1996)
+\bibitem{2_Piyavskij}	Piyavskij, S.A.: An Algorithm for Finding the Absolute Extremum of a Function. USSR Comput. Math. Math. Phys. \textbf{12}(4), 57--67 (1972)
+\bibitem{2_SergDivBest} Sergeyev, Y.D.: On convergence of “Divide the Best” global optimization algorithms. Optimization 44(3), 303--325 (1998)
+\bibitem{2_SergKvasMukhOZ} Sergeyev, Y.D., Kvasov, D.E., Mukhametzhanov, M.S.: Operational zones for comparing metaheuristic and deterministic one-dimensional global optimization algorithms.
+Mathematics and Computers in Simulation \textbf{141}, 96-109 (2017)
+\bibitem{2_SergStrAGSIP} Sergeyev, Y.D., Strongin, R.G.:  A global minimization algorithm with parallel iterations. Comput. Maths. Math. Phys. \textbf{29}(2), 7-15 (1989)
+\bibitem{2_SergStrLeraMonogr}	Sergeyev, Y.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting Space-Filling Curves. Springer (2013)
+\bibitem{2_ShiOlaf}   Shi, L., {\'O}lafsson, S.: Nested Partitions Method for Global Optimization. Operations Research \textbf{48}, 390--407 (2000)
+\bibitem{2_Shubert} Shubert, B.O.: A sequential method seeking the global maximum of a function. SIAM J. Numer. Anal. \textbf{9}(3), 379--388 (1972)
+\bibitem{2_StrMonRus}	Strongin, R.G.: Numerical Methods in Multiextremal Problems (Information-Statistical Algorithms). Nauka, Moscow (1978). In Russian
+\bibitem{2_Str1989} Strongin, R.G.: The information approach to multiextremal optimization problems. Stochastics and Stochastic Reports \textbf{27}, 65-82 (1989).
+\bibitem{2_StrSergMon2000} Strongin, R.G., Sergeyev, Y.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000)
+
+\end{thebibliography}
 %\end{document}

### 2017-11-06 2017/ParOptBook/Chapter4.tex
Chapter 2 included

@@
 
-\section*{A brief conclusion}
-Constrained global optimization problems are under consideration.
-The method of penalty functions is one of the most popular numerical methods for solving the problems of this kind.
-Nevertheless, the method of penalty functions has a number of disadvantages.
-A new method for reduction of a constrained problem to an unconstrained one, so called \emph{index method}, has been considered in this chapter.
-This method, unlike the classical penalty function one, allows avoiding the problem of the parameter selection (like the penalty coefficients for the constraints).
-It is also allows solving the problems with partially defined objective function and constraints.
-This is quite natural for many applied problems, especially, in the optimal design of technical systems, because if some conditions of operation are not met, then some other characteristics of the system performance may not be defined.
-
-The issues of convergence acceleration of the index method have been considered also.
-Two approaches for the convergence acceleration have been proposed.
-The first of them is based on concept of $\varepsilon$-reserved solution to the constrained problem.
-The role of $\varepsilon$-reserved solution in constrained global optimization problems is somewhat similar to the regularity requirements in classical nonlinear programming problems.
-The second approach is based on adaptive order of checking for constraints.
-In the index method, every iteration performed at the corresponding point of the search domain includes checking for constraints of the problem at this point.
-When a violation is discovered for the first time, the current iteration is stopped, and the next iteration is initiated.
-However, in Lipschitz optimization problems a constraint violated at a point of the search domain is also violated in a certain neighborhood of this point.
-In this case, to reduce the number of checks, it is expedient to begin iterations in the neighborhood of this point with checking for the violated constraint.
-Thus, the iteration can be completed at a lower cost.
-
-For all proposed algorithms, the results of numerical experiments are presented. The results confirm the theoretical statements about convergence acceleration.
 \begin{thebibliography}{99.}

### 2017-11-07 2018/ITNT/ITNT2018.tex
С фрагментами русского текста

@@
 \begin{abstract}
-All articles must contain an abstract.
-The abstract should give readers concise information about the content of the article and indicate the main results obtained and conclusions drawn.
-The abstract is not part of the text and should be complete in itself; no table numbers, figure numbers, references or displayed mathematical expressions should be included.
-It should be suitable for direct inclusion in abstracting services and should not normally exceed 200 words in a single paragraph.
-Since contemporary information-retrieval systems rely heavily on the content of titles and abstracts to identify relevant articles in literature searches, great care should be taken in constructing both.
+\Russian
+В работе рассматриваются задачи многоэкстремальной оптимизации и высокопроизводительный параллельный алгоритм их решения.
+Исследование масштабируемости алгоритма проводится на классе задач, в которых трудоемкость вычисления функций зависит от точки проведения итерации.
+Предложенный в работе алгоритм может задействовать как CPU (для решения более сложных подзадач), так и GPU (для решения простых подзадач).
+Приведены результаты вычислительных экспериментов, демонстрирующие ускорение при решении серии multiextremal con-strained problems.
 \end{abstract}
@@
 \section{Introduction}
+\Russian
+В статье рассматриваются задачи многоэкстремальной оптимизации с ограничениями и параллельные методы их решения.
+Важной особенностью многоэкстремальных задач является тот факт, что глобальный экстремум есть интегральная характеристика задачи, таким образом, его отыскание связано с построением покрытия области поиска и вычислением значений оптимизируемой функции во всех точках этого покрытия.
+На сложность решения задач рассматриваемого класса решающее влияние оказывает размерность: вычислительные затраты растут экспоненциально при ее увеличении, поэтому для решения подобных задач требуются методы, которые порождают в области поиска существенно неравномерную сетку, более плотную в окрестности глобального минимума и разреженную вдали от него.
+В данной статье изложены результаты исследования масштабирования параллельного индексного алгоритма глобальной оптимизации, разработанного в ННГУ им.
+Н.И.
+Лобачевского [1, 2] и реализованного в системе Globalizer [NACO].
 
-This file should be used as a pattern for English Thesis. You can write on email mingazov88@gmail.com with questions concerning the preparation of the English Thesis.
+В рамках обсуждаемого подхода решение многомерных задач сводится к решению набора связанных подзадач меньшей размерности.
+Соответствующая редукция основана на использовании разверток единичного отрезка вещественной оси на гиперкуб.
+Роль таких разверток играют непрерывные однозначные отображения типа кривой Пеано, называемые также кривыми, заполняющими пространство.
+Еще одним используемым механизмом снижения размерности решаемой задачи является nested optimization scheme.
+Численные методы, позволяющие эффективно использовать аппарат таких отображений, детально разработаны и обоснованы в [1, 2].
+Алгоритмы, развиваемые в данной статье, основаны на предположении липшицевости оптимизируемого критерия и ограничений задачи, что является типичными для многих других подходов к построению параллельных оптимизационных алгоритмов (см., например, [3 – 5).
+Предположение такого рода выполняется для многих прикладных задач, поскольку относительные вариации функций, характеризующих моделируемую систему, обычно не превышают некоторый порог, определяемый ограниченной энергией изменений в системе.
+Также стандартным является предположение, что проведения одного испытания (вычисления значений целевой функции и функций ограничений в точке из области поиска) является трудоемкой операцией, т.к.
+использует результаты численного моделирования.
+
+Новым элементом, исследуемым в данной работе, является предположение о разной трудоемкости вычисления функций задачи в зависимости от разных компонент вектора параметров.
+Предполагается, что в функциях можно выделить связанные между собой ``сложную'' и ``легкую'' части.
+Тогда решение ``сложной'' части задачи (для которой требуется реализация трудоемких вычислительных алгоритмов для проведения испытаний) можно провести на CPU, использовав экономный индексный алгоритм глобальной оптимизации.
+``Легкую'' же часть задачи (реализация которой не требует сложных алгоритмов и легко портируется на ускоритель) можно решить на GPU с использованием переборных алгоритмов.
+Очевидно, что при этом на GPU будет проведено значительно больше испытаний, чем на CPU, но из-за различия в сложности подзадач, решаемых на разных устройствах, ожидается ускорение алгоритма в целом.
+Предложенный подход реализован в разработанной в ННГУ им. Н.И. Лобачевского параллельной программной системе решения задач глобальной оптимизации Globalizer.
 
@@
 \Russian
-ѕараллельный индексный алгоритм, который может быть применен дл€ решени€ подобных задач с частично вычислимыми функци€ми и который реализован в системе Globalizer, подробно описан в [2].
-ќсновна€ иде€ алгоритма состоит в сведении исходной многомерной задачи к набору св€занных подзадач меньшей размерности, и параллельному их решению.
-—хемы редукции размерности, используемые при работе индексного алгоритма, кратко описаны в следующем разделе.
+Параллельный индексный алгоритм, который может быть применен для решения подобных задач с частично вычислимыми функциями и который реализован в системе Globalizer, подробно описан в [2].
+Основная идея алгоритма состоит в сведении исходной многомерной задачи к набору связанных подзадач меньшей размерности, и параллельному их решению.
+Схемы редукции размерности, используемые при работе индексного алгоритма, кратко описаны в следующем разделе.
 
@@
 \Russian 
-¬ проведенном исследовании данна€ схема примен€лась при M=2, т.е. использовалс€ только один уровень вложенности
+В проведенном исследовании данная схема применялась при $M=2$, т.е. использовался только один уровень вложенности
 \begin{equation}
@@
 
+\Russian
+Организация параллельных вычислений с использованием рекурсивной схемы оптимизации была подробно рассмотрена для общей/распределенной памяти, а также ускорителей в [6, 7].
+Однако в данных работах были рассмотрены задачи, время проведения испытаний в которых не зависит от точки проведения испытания.
+Здесь же задача рассматривается в предположении, что в функции $\varphi(y_1,...,y_N)$ можно выделать более сложную часть $f(y_1,...,y_N)$ (зависящую от всех параметров задач) и более простую часть $g(y_s,...,y_N )$ (зависящую лишь от части параметров), например, $\varphi(y)= f(y)g(y)$.
+
+Сложная часть функции подразумевает проведение численных расчетов, связанных с численным моделированием, которые могут быть выполнены только на CPU.
+Простая часть не предполагает сложных расчетов и может быть вычислена на ускорителе, например, GPU.
+Для решения задачи подобной структуры можно использовать параллельную схему рекурсивной оптимизации с использованием CPU на верхнем и GPU на нижнем уровне рекурсии.
+
 \section{Numerical experiments}\label{sec:5}
+
+\Russian
+Рекурсивная схема решения global optimization problems реализована в решателе Globzlizer, разработанном в ННГУ им.
+Н.И.
+Лобачевского.
+Алгоритмическую основу Globalizer составляют методы глобального поиска и различные схемы редукции размерности.
+Численные эксперименты, результаты которых отражены в [9,10], показывают, что данные методы, как минимум, не уступают известным методам аналогичного назначения, а по некоторым параметрами и превосходят их.
+
+Исследование масштабируемости рекурсивного алгоритма будем проводить, решая серию из 100 тестовых задач условной глобальной оптимизации.
+В работе [11] предложен подход, который позволяют формировать задачи условной глобальной оптимизации со следующими свойствами:
 
@@
 \end{itemize}
+
+\Russian
+С целью имитации прикладных задач с разной вычислительной сложностью мы будем использовать комбинацию классов функций вида 
+\[
+\varphi(y_1,...,y_N) = p(y_1,...,y_N)(f(y_1,y_2)+g(y_3,...,y_N))
+\]
+Здесь $f(y_1,y_2)$ - двумерная функция из класса, описанного в [11], $g(y_3,...,y_N)$ - функция размерности $N-2$ из классов, описанных в [12], $p(y_1,...,y_N)$ – полином второй степени.
+Домножение на $p(y_1,...,y_N)$ исключает возможность сепарабельного поиска минимумов у функций $f(y_1,y_2)$ и $g(y_3,...,y_N)$.
+
+Линии уровня двумерных подзадач на основе функций  $f(y_1,y_2)$ и $g(y_3,y_4)$ приведены на Fig.~\ref{fig_1}.
+Видно, что подзадачи (a) обладают более сложной структурой по сравнению с подзадачей (b).
+одновременно с этим вычисление значения функции $f(y_1,y_2)$ является более трудоемким, чем $g(y_3,y_4)$.
 
@@
 \end{minipage}
-\center{{\bf Figure 1.} The level lines of (a) hard and (b) simple subproblems }
+\center{{\bf Figure 1.} The level lines of (a) hard and (b) simple subproblems}
+\label{fig_1}
 \end{figure}
@@
 \Russian
-ѕервый эксперимент проведем при решении серий задач Simple и Hard на одном узле, полностью задействовать оба доступных CPU (т.е.
-p=16 физических €дер).
-¬ таблице 1 приведено среднее врем€ (в секундах), потребовавшихс€ дл€ решени€ задач серии.
+Первый эксперимент проведем при решении серий задач Simple и Hard на одном узле, полностью задействовать оба доступных CPU (т.е.
+p=16 физических ядер).
+В таблице 1 приведено среднее время (в секундах), потребовавшихся для решения задач серии.
 
@@
 \Russian
-—ледующий эксперимент проведем, задействовав p узлов по три графических ускорител€ на каждом из них.
-“ак, при p=32 было задействовано 94 GPU ускорител€, на каждом из которых Ц 2688 CUDA-€дер.
-¬ таблице 2 приведено врем€ работы алгоритма, а в таблице 3 Ц ускорение по отношению к запуску на одном узле.
+Следующий эксперимент проведем, задействовав p узлов по три графических ускорителя на каждом из них.
+Так, при p=32 было задействовано 94 GPU ускорителя, на каждом из которых - 2688 CUDA-ядер.
+В таблице 2 приведено время работы алгоритма, а в таблице 3 - ускорение по отношению к запуску на одном узле.
 
@@
 Computational experiments were carried out on a high-performance cluster of Lobachevsky State University of Nizhni Novgorod.
-The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM, and three NVIDIA Kepler K20’ GPUs (2688 CUDA cores, 6 Gb GDDR5).
+The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM, and three NVIDIA Kepler K20Х GPUs (2688 CUDA cores, 6 Gb GDDR5).
 
 \section{Conclusion}
+\Russian
+Результаты проведенных экспериментов на серии тестовых задач условной глобальной оптимизации с разным временем вычисления функций задачи показывают, что индексный алгоритм глобальной оптимизации в сочетании с блочной рекурсивной схемой редукции размерности показывают хорошее ускорение на рассматриваемом классе задач.
+Направлением дальнейших исследований является расширение круга алгоритмов, используемых для решения подзадач на GPU.
 

### 2017-11-09 2018/ITNT/ITNT2018.tex
Translate to English

--- 2017-11-03 2018/ITNT/ITNT2018.tex
@@
 \usepackage{graphicx}
-\usepackage[russian,english]{babel}
+\usepackage[english]{babel}
 \usepackage{amsmath}
@@
 \begin{document}
-\title{High Performance Computing for Global Optimization Problems}
+\title{High performance computing for global optimization problems}
 
@@
 
-\address{Lobachevsky State University of Nizhni Novgorod, Gagarin avenue 23, Nizhni Novgorod, Russia, 603950}
+\address{Lobachevsky State University of Nizhni Novgorod, Gagarin avenue 23, 
+Nizhni Novgorod, Russia, 603950}
 
 \begin{abstract}
-All articles must contain an abstract.
-The abstract should give readers concise information about the content of the article and indicate the main results obtained and conclusions drawn.
-The abstract is not part of the text and should be complete in itself; no table numbers, figure numbers, references or displayed mathematical expressions should be included.
-It should be suitable for direct inclusion in abstracting services and should not normally exceed 200 words in a single paragraph.
-Since contemporary information-retrieval systems rely heavily on the content of titles and abstracts to identify relevant articles in literature searches, great care should be taken in constructing both.
+In the present work, the multiextremal optimization problems and a high-
+performance parallel algorithm for solving these ones are considered. The 
+investigation of the algorithm scalability has been carried out on the 
+problem class, in which the computation costs of the functions depended on 
+the iteration point. The algorithm proposed in the present work can utilize 
+the CPUs (for solving more complex subproblems) as well as the GPUs (for 
+solving the simple subproblems). The results of numerical experiments 
+demonstrating the speedup when solving a series of multiextremal constrained 
+problems are presented.
 \end{abstract}
@@
 \section{Introduction}
+In the present paper, the constrained multiextremal optimization problems and 
+the parallel methods for solving these ones are considered. The fact that the 
+global extremum is an integral characteristic of a problem is an important 
+feature of the multiextremal problems. Thus, its finding is related to the 
+construction of a coverage of the search domain and the computing the 
+optimized function values at all points of this coverage. The dimensionality 
+affects the difficulty of solving the problems of considered class crucially: 
+the computational costs grow with increasing this one exponentially, 
+therefore, for solving such problems, the methods are required, which 
+generate an essentially nonuniform grid in the search domain, denser in the 
+nearness of the global minimum and more sparse far away from this one. In the 
+present paper, the results of investigation of the scaling of the parallel 
+index global optimization algorithm developed in Lobachevsky State 
+University of Nizhni Novgorod \cite{Strongin2000,Strongin2013} and implemented in the Globalizer 
+system [NACO] are presented.
 
-This file should be used as a pattern for English Thesis. You can write on email mingazov88@gmail.com with questions concerning the preparation of the English Thesis.
+Within the framework of the discussed approach, the solving of the 
+multidimensional problems is reduced to the solving of a set of the connected 
+subproblems of less dimensionality. The corresponding dimensionality 
+reduction is based on the use of the evolvents of a unit interval on the real 
+axis onto a hypercube. The continuous unambiguous mappings like Peano curves, 
+called also the space-filling curves, play the role of such evolvents. 
+The nested optimization scheme is one more mechanism used for reducing the 
+dimensionality of the problem being solved. The numerical methods allowing 
+the efficient utilization of such reduction schemes had been developed 
+in details and substantiated in \cite{Strongin2000,Strongin2013}.
+
+In the algorithms developed in the present paper the objective function and constraints are assumed to be Lipschitzian that is typical for many other approaches to the development of the parallel optimization algorithms (see, for example, \cite{Jones,Zilinskas,Evtushenko}).
+This assumption is natural for many applied problems since the relative variations of the functions generally cannot exceed a certain threshold determined by the bounded energy of changes occurring in the system under study.
+Also, a standard assumption is that the execution of a single trial (the computing of the values of the objective function and constraints at a point in the search domain) is a time-consuming operation since it
+utilized the results of numerical modeling.
+
+An assumption on different costs of computing the problem functions subject to 
+various components of the vector of parameters is a novel element investigated in the present work. 
+It is assumed that it is possible to select the ``difficult'' and ``easy'' parts 
+of the functions connected with each other. Then, the solving of the 
+``difficult'' part of the problem (for which the execution of the computation-
+costly algorithms is required to conduct the trials) can be performed on the 
+CPU using an efficient index global optimization algorithm. The ``easy'' part 
+of the problem (the execution of which doesn’t require the complex algorithms 
+and can be transferred to an accelerator easily) can be solved on a GPU using 
+the uniform grid technique. Obviously, in this case a major portion of the trials 
+will be performed on the GPUs, and the minor one -- on the CPU. However, 
+because of the difference in the difficulties of the subproblems solved on 
+different units, a speedup of the algorithm as a whole can be expected. 
+The proposed approach has been implemented in Globalizer parallel software 
+system for solving the global optimization problem developed in Lobachevsky 
+State University of Nizhni Novgorod.
 
@@
 \begin{equation}\label{problem}
-\varphi(y^\ast)=\min{\left\{\varphi(y): y \in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}
+\varphi(y^\ast)=\min{\left\{\varphi(y): y \in D, \; g_i(y)\leq 0, \; 1 \leq i 
+\leq m\right\}}
 \end{equation}
@@
 m+1$, respectively, and may be multiextremal. It is assumed that the
-functions $g_i(y),\; 1 \leq i \leq m+1$, are defined and computable only in the corresponding domains
+functions $g_i(y),\; 1 \leq i \leq m+1$, are defined and computable only in 
+the corresponding domains
 \[
-Q_1=D, \; Q_{i+1}=\left\{y \in Q_i : g_i(y) \leq 0 \right\}, \; 1 \leq i \leq m.
+Q_1=D, \; Q_{i+1}=\left\{y \in Q_i : g_i(y) \leq 0 \right\}, \; 1 \leq i \leq 
+m.
 \]
@@
 computed at this point.
-%отсюда
 
 Thus, a \textit{trial} at a point $y^k \in D$ executed at the $k$-th
-iteration of the algorithm will consist in computing the values $g_1(y),...,g_\nu(y),$ where the
+iteration of the algorithm will consist in computing the values 
+$g_1(y),...,g_\nu(y),$ where the
 index $\nu \leq m$ is determined by the conditions
@@
 
-\Russian
-ѕараллельный индексный алгоритм, который может быть применен дл€ решени€ подобных задач с частично вычислимыми функци€ми и который реализован в системе Globalizer, подробно описан в [2].
-ќсновна€ иде€ алгоритма состоит в сведении исходной многомерной задачи к набору св€занных подзадач меньшей размерности, и параллельному их решению.
-—хемы редукции размерности, используемые при работе индексного алгоритма, кратко описаны в следующем разделе.
+The parallel index algorithm, which can be applied to solving such problems 
+with partially defined functions and which has been implemented in the 
+Globalizer system, described in details in \cite{Strongin2013}. The main idea of the algorithm 
+consists in the reduction of the initial multidimensional problem to a set of 
+the nested subproblems of less dimensionality and the solving of these ones 
+in parallel. The dimensionality reduction schemes utilized in the operation 
+of the index algorithm are described briefly in the next section.
 
@@
 \begin{equation}
-\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq N\right\}=\left\{y(x):0\leq x \leq 1 \right\}
+\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq 
+N\right\}=\left\{y(x):0\leq x \leq 1 \right\}
 \end{equation}
-unambiguously mapping the interval of real axis $[0,1]$ onto a $N$-dimensional cube is the first of the dimensionality reduction methods considered.
-To implement this method of dimensionality reduction a numerically constructed curve (\textit{evolvent}) is used.
-The evolvent is $2^{-m}$ accurate approximation of the theoretical Peano curve in $L_\infty$ metric, where $m$ is an evolvent construction parameter.
-Problems of numerical construction of the evolvents and the corresponding theory are considered in detail in \cite{Strongin2000}.
+unambiguously mapping the interval of real axis $[0,1]$ onto a $N$-
+dimensional cube is the first of the dimensionality reduction methods 
+considered.  To implement this method of dimensionality reduction a 
+numerically constructed curve (\textit{evolvent}) is used. The evolvent is 
+$2^{-m}$ accurate approximation of the theoretical Peano curve in $L_\infty$ 
+metric, where $m$ is an evolvent construction parameter. Problems of 
+numerical construction of the evolvents and the corresponding theory are 
+considered in detail in \cite{Strongin2000}.
 
-By using this kind of mapping it is possible to reduce the multidimensional problem~(\ref{problem}) to a univariate problem
+By using this kind of mapping it is possible to reduce the multidimensional 
+problem~(\ref{problem}) to a univariate problem
 \[
-\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
+\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 
+0, \; 1 \leq i \leq m\right\}.
 \]
@@
 \[
-\left|g_i(y(x'))-g_i (y(x''))\right| \leq H_i \left|x'-x'' \right|^{1/N}, \; x',x''\in [0,1], \; 1\leq i \leq m+1.
+\left|g_i(y(x'))-g_i (y(x''))\right| \leq H_i \left|x'-x'' \right|^{1/N}, \; 
+x',x''\in [0,1], \; 1\leq i \leq m+1.
 \]
@@
 
-Nested optimization scheme is based on relation (see \cite{Grishagin2001})
+Nested optimization scheme is based on relation (see \cite{Strongin2013})
 \begin{equation}\label{nested}
-\min_{y \in D}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}= \min_{u_1\in D_1}\min_{u_2\in D_2}...\min_{u_M\in D_M }{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
+\min_{y \in D}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq 
+m\right\}}= \min_{u_1\in D_1}\min_{u_2\in D_2}...\min_{u_M\in D_M 
+}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
 \end{equation}
-which allows replacing the solving of multidimensional problem (\ref{problem}) by solving a family of subproblems related to each other recursively.
+which allows replacing the solving of multidimensional problem 
+(\ref{problem}) by solving a family of subproblems related to each other 
+recursively.
 Here we consider vector $y$ as a vector of block variables
@@
 \end{equation}
-where the $i$-th block variable $u_i$ is a vector of $N_i$ components of vector $y$, taken serially i.e.
-$u_1=(y_1,y_2,...,y_{N_1})$, $u_2=(y_{N_1+1},y_{N_1+2},...,y_{N_1+N_2})$,..., $u_M=(y_{N-N_M+1},y_{N-N_M+2},...,y_{N})$, at that $N_1+N_2+...+N_M=N$.
-The subdomains $D_i, 1 \leq i \leq M$, are projections of initial search domain $D$ onto the subspaces corresponding to the variables $u_i, 1 \leq i \leq M$.
+where the $i$-th block variable $u_i$ is a vector of $N_i$ components of 
+vector $y$, taken serially i.e. $u_1=(y_1,y_2,...,y_{N_1})$, 
+$u_2=(y_{N_1+1},y_{N_1+2},...,y_{N_1+N_2})$,..., $u_M=(y_{N-N_M+1},y_{N-
+N_M+2},...,y_{N})$, at that $N_1+N_2+...+N_M=N$. 
+The subdomains $D_i, 1 \leq i \leq M$, are projections of initial search 
+domain $D$ onto the subspaces corresponding to the variables $u_i, 1 \leq i 
+\leq M$.
 
-\Russian 
-¬ проведенном исследовании данна€ схема примен€лась при M=2, т.е. использовалс€ только один уровень вложенности
+In the present study, this scheme has been applied for $M=2$, i.e. only one 
+nesting level has been used
 \begin{equation}
-\min_{y \in D}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}= \min_{u_1\in D_1}\min_{u_2\in D_2}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}.
+\min_{y \in D}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq 
+m\right\}}= \min_{u_1\in D_1}\min_{u_2\in D_2}{\left\{\varphi(y): \; 
+g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}.
 \end{equation}
 
-\section{Organization of Parallel Computing}\label{sec:4}
+\section{Organization of parallel computing}\label{sec:4}
+
+The organization of the parallel computations with the use of the recursive 
+optimization scheme has been considered in details for the shared/distributed 
+memory as well as for the accelerators in \cite{Sysoyev,BarkalovLebedev}. However, in these works the 
+problems were considered, in which the time of the trial execution didn’t 
+depend on the trial point. Here, the problem is considered assuming that in 
+the function $\varphi(y_1,...,y_N)$ it is possible to select more difficult 
+part $f(y_1,...,y_N)$ (depending on all problem parameters) and a simpler 
+part $g(y_s,...,y_N )$ (depending on a part of the parameters only), for 
+example, $\varphi(y)= f(y)g(y)$.
+
+The difficult part of the function implies performing some  
+computations related to the numerical simulation, which can be performed on the 
+CPUs only. The simple part doesn’t imply the complex computations and can be 
+computed on an accelerator, for example, GPU. For solving a problem with such 
+a structure, one can apply the parallel recursive optimization scheme with 
+the use the CPUs at the upper nesting level and the GPUs at the lower one.
 
@@
 
+The recursive scheme of solving the global optimization problems has been 
+implemented in the Globzlizer solver developed in Lobachevsky State 
+University of Nizhni Novgorod. The global search methods and various 
+dimensionality reduction schemes make the algorithmic base for the 
+Globalizer. The numerical experiments, the results of which are presented in 
+\cite{BarkalovGergelLebedev,BarkalovGergel} demonstrate these methods, at least, are not worse than the well known 
+methods applied for similar purposes, and even overcome these ones with 
+respect to some parameters.
+
+Let us conduct the investigation of the scalability of the parallel 
+algorithm by solving a series of 100 test problems of the constrained global 
+optimization. In work \cite{Gergel2017} the approach has been proposed, which allowed 
+generating the constrained global optimization problems with the following 
+properties:
 \begin{itemize}
-	\item one could control the size of the feasible domain $Q_{m+1}$ with respect to the whole domain $D$;
-	\item the global minimizer of the objective function is known a priori taking into account the constraints;
-	\item the global minimizer of the objective function without accounting for the constraints is out of the feasible domain $Q_{m+1}$  (with the purpose of simulating the behavior of the constraints and the objective function in the applied constrained optimization problems).
+	\item one could control the size of the feasible domain $Q_{m+1}$ with 
+respect to the whole domain $D$;
+	\item the global minimizer of the objective function is known a priori 
+taking into account the constraints;
+	\item the global minimizer of the objective function without accounting 
+for the constraints is out of the feasible domain $Q_{m+1}$  (with the 
+purpose of simulating the behavior of the constraints and the objective 
+function in the applied constrained optimization problems).
 \end{itemize}
+
+In order to simulate the applied problems with various computational 
+difficulty, we will use a combination of the function classes of the kind 
+\[
+\varphi(y_1,...,y_N) = p(y_1,...,y_N)(f(y_1,y_2)+g(y_3,...,y_N)).
+\]
+Here $f(y_1,y_2)$ is a two-dimensional function from the class described in \cite{Gergel2016}, $g(y_3,...,y_N)$ is a function of the dimensionality $N-2$ from the 
+class described in \cite{Sergeyev2015}, and $p(y_1,...,y_N)$ is a second 
+order polynomial. The multiplication by $p(y_1,...,y_N)$ excluded the possibility of 
+separable search of the minima of the functions $f(y_1,y_2)$ and 
+$g(y_3,...,y_N)$.
+
+The level lines of the two-dimensional subproblems based on the functions 
+$f(y_1,y_2)$ and $g(y_3,y_4)$ are shown in figure~1. The feasible domains are highlighted by color. One can see that 
+the subproblem (a) has more complex structure as compared to the subproblem 
+(b). At the same time, computing the values of the function $f(y_1,y_2)$ is 
+more computation-costly than of $g(y_3,y_4)$.
 
@@
 \end{minipage}
-\center{{\bf Figure 1.} The level lines of (a) hard and (b) simple subproblems }
+\center{{\bf Figure 1.} The level lines of (a) hard and (b) simple 
+subproblems}
 \end{figure}
 
-The numerical experiments were carried out using two classes of problems (\textit{Simple} ans \textit{Hard}) with $N=5$.
-The problem was considered to be solved if the algorithm generates a trial point $y^k$ in $\delta$-vicinity of the global minimizer, i.e.
-$\left\|y^k-y^\ast\right\|\leq\delta$ .
-The size of the vicinity was selected as $\delta=0.01\left\|b-a\right\|$, where $a$ and $b$ are the boundaries of the search domain $D$.
-The maximum allowable number of iterations was $K_{max} = 10^6$.
+The numerical experiments were carried out using two classes of problems 
+(\textit{Simple} ans \textit{Hard}) with $N=5$. The problem was considered to 
+be solved if the algorithm generates a trial point $y^k$ in $\delta$-vicinity 
+of the global minimizer, i.e. $\left\|y^k-y^\ast\right\|\leq\delta$ . The 
+size of the vicinity was selected as $\delta=0.01\left\|b-a\right\|$, where 
+$a$ and $b$ are the boundaries of the search domain $D$. The maximum 
+allowable number of iterations was $K_{max} = 10^6$.
 
-\Russian
-ѕервый эксперимент проведем при решении серий задач Simple и Hard на одном узле, полностью задействовать оба доступных CPU (т.е.
-p=16 физических €дер).
-¬ таблице 1 приведено среднее врем€ (в секундах), потребовавшихс€ дл€ решени€ задач серии.
+Let us conduct the first experiment by solving the \textit{Simple} and \textit{Hard} problem 
+series on a single node employing both available CPUs in full (i. e. $p=16$ 
+cores available). In table 1 the averaged time (in seconds) 
+required to solve the problems of the series is presented.
 
@@
 \hline
-Problems & Time, sec. \\
+Problems & Time (sec) \\
 \hline
@@
 
-\Russian
-—ледующий эксперимент проведем, задействовав p узлов по три графических ускорител€ на каждом из них.
-“ак, при p=32 было задействовано 94 GPU ускорител€, на каждом из которых Ц 2688 CUDA-€дер.
-¬ таблице 2 приведено врем€ работы алгоритма, а в таблице 3 Ц ускорение по отношению к запуску на одном узле.
+Let us conduct the next experiment employing $p$ nodes each including three 
+graphic accelerators. Thus, at $p=32$ total 96 GPU accelerators were employed 
+each including 2688 CUDA cores. In table 2, the execution time of the 
+algorithm is presented and in table 3 -- the speedup with respect to the run 
+on a single node.
 
@@
 
-Computational experiments were carried out on a high-performance cluster of Lobachevsky State University of Nizhni Novgorod.
-The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM, and three NVIDIA Kepler K20’ GPUs (2688 CUDA cores, 6 Gb GDDR5).
+Computational experiments were carried out on a high-performance cluster of 
+Lobachevsky State University of Nizhni Novgorod. The cluster node included 
+two Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM, and three NVIDIA 
+Kepler K20Õ GPUs (2688 CUDA cores, 6 Gb GDDR5). 
+The results of experiments demonstrate that the paralle index algorithm combined with the nested optimization scheme provides a good speedup on the considered problem class.
 
-\section{Conclusion}
+%The results of experiments performed on a series of test problems of the constrained global optimization with different time of computing the the problem functions demonstrate the index global optimization algorithm combined with the block recursive scheme for the dimensionality reduction to provide a good speedup on the considered problem class.
+The directions of further investigations include the expansion of the number of algorithms used for solving the subproblems on the GPUs.
 
 \section*{Acknowledgments}
-This study was supported by the Russian Science Foundation, project No 16-11-10150.
+This study was supported by the Russian Science Foundation, project No 16-11-
+10150.
 
@@
 \bibitem{Strongin2000}
-Strongin, R.G. Global optimization with non-convex constraints. Sequential and parallel algorithms. / R.G. Strongin, Ya.D.  Sergeyev --- Dordrecht: Kluwer Academic Publishers, 2000. --- 704 p.
+Strongin, R.G. Global optimization with non-convex constraints. Sequential 
+and parallel algorithms. / R.G. Strongin, Ya.D.  Sergeyev --- Dordrecht: 
+Kluwer Academic Publishers, 2000. --- 704 p.
 
 \bibitem{Strongin2013}
-Strongin, R.G.
-Parallel computing in global optimization problems / R.G.
-Strongin, V.P.
-Gergel, V.A.
-Grishagin, K.A.
-Barkalov --- Moscow: Publishing of the Moscow State University, 2013.
---- 280 p.
---- (in Russian).
+Strongin, R.G. Parallel computing in global optimization problems / R.G. 
+Strongin, V.P. Gergel, V.A. Grishagin, K.A. Barkalov --- Moscow: Publishing 
+of the Moscow State University, 2013. --- 280 p. --- (in Russian).
 
 \bibitem{Jones}
-Jones, D.R. The DIRECT global optimization algorithm / In: Floudas, C. A., Pardalos, P. M. (eds.) The Encyclopedia of Optimization, Second Edition. --- Heidelberg: Springer, 2009. --- P. 725--735.
+Jones, D.R. The DIRECT global optimization algorithm / In: Floudas, C. A., 
+Pardalos, P. M. (eds.) The Encyclopedia of Optimization, Second Edition. --- 
+Heidelberg: Springer, 2009. --- P. 725--735.
 
-\bibitem{Zilinskas2011}
-Paulavi\v{c}ius, R.
-Parallel branch and bound for global optimization with combination of Lipschitz bounds / R.
-Paulavi\v{c}ius, J.
-\v{Z}ilinskas, A.
-Grothey // Optimization Methods \& Software.
---- 2011.
---- Vol.
-26(3).
---- P.
+\bibitem{Zilinskas}
+Paulavi\v{c}ius, R. Parallel branch and bound for global optimization with 
+combination of Lipschitz bounds / R. Paulavi\v{c}ius, J. \v{Z}ilinskas, A. 
+Grothey // Optimization Methods \& Software. --- 2011. --- Vol. 26(3). --- P. 
 487--498.
@@
 \bibitem{Evtushenko}
-Evtushenko, Yu.G.
-Parallel global optimization of functions of several variables / Yu.G.
-Evtushenko, V.U.
-Malkova, A.A.
-Stanevichyus // Computational Mathematics and Mathematical Physics.
---- 2009.
---- Vol.
-49 (2).
---- P.
-246--260.
+Evtushenko, Yu.G. Parallel global optimization of functions of several 
+variables / Yu.G. Evtushenko, V.U. Malkova, A.A. Stanevichyus // 
+Computational Mathematics and Mathematical Physics. --- 2009. --- Vol. 49 
+(2). --- P. 246--260.
 
 \bibitem{Sysoyev}
-Sysoyev, A.V.
-MPI implementation of dimension reduction multilevel scheme for parallel solving the global optimization problems / A.V.
-Sysoyev, K.A.
-Barkalov, V.P.
-Gergel, I.G.
-Lebedev // Russian Supercomputing Days: Proceedings of the International Scientific Conference.
---- Moscow: Publishing of the Moscow State University, 2015.
---- P.
-61--68.
+Sysoyev, A.V. MPI implementation of dimension reduction multilevel scheme for 
+parallel solving the global optimization problems / A.V. Sysoyev, K.A. 
+Barkalov, V.P. Gergel, I.G. Lebedev // Russian Supercomputing Days: 
+Proceedings of the International Scientific Conference. --- Moscow: 
+Publishing of the Moscow State University, 2015. --- P. 61--68.
 
 \bibitem{BarkalovLebedev}
-Barkalov, K.A.
-Solving global optimization problems on GPU / K.A.
-Barkalov, I.G.
-Lebedev // Russian Supercomputing Days: Proceedings of the international conference.
---- Moscow: Publishing of the Moscow State University, 2016.
---- P.
-640--650.
+Barkalov, K.A. Solving global optimization problems on GPU / K.A. Barkalov, 
+I.G. Lebedev // Russian Supercomputing Days: Proceedings of the international 
+conference. --- Moscow: Publishing of the Moscow State University, 2016. --- 
+P. 640--650.
 
 \bibitem{BarkalovGergel}
-Barkalov, K. Parallel global optimization on GPU / K. Barkalov K., V. Gergel // Journal of Global Optimization. --- 2016. --- Vol. 66(1). --- P. 3--20.
+Barkalov, K. Parallel global optimization on GPU / K. Barkalov K., V. Gergel 
+// Journal of Global Optimization. --- 2016. --- Vol. 66(1). --- P. 3--20.
 
 \bibitem{BarkalovGergelLebedev}
-Barkalov, K. Use of Xeon Phi Coprocessor for Solving Global Optimization Problems / K. Barkalov, V. Gergel, I. Lebedev //  Lecture Notes in Computer Science. --- 2015. --- Vol. 9251. --- P. 307--318.
+Barkalov, K. Use of Xeon Phi Coprocessor for Solving Global Optimization 
+Problems / K. Barkalov, V. Gergel, I. Lebedev //  Lecture Notes in Computer 
+Science. --- 2015. --- Vol. 9251. --- P. 307--318.
 
 \bibitem{Gergel2017}
-Gergel, V. An Approach for Generating Test Problems of Constrained Global Optimization / V. Gergel // Lecture Notes in Computer Science. --- 2017. --- Vol. 10556. --- P. 314--319.
+Gergel, V. An Approach for Generating Test Problems of Constrained Global 
+Optimization / V. Gergel // Lecture Notes in Computer Science. --- 2017. --- 
+Vol. 10556. --- P. 314--319.
 
 \bibitem{Gergel2016}
-Gergel, V. Adaptive nested optimization scheme for multidimensional global search / V. Gergel, V. Grishagin, A. Gergel // Journal of Global Optimization. --- 2015. --- Vol. 66(1). --- P. 35--51.
+Gergel, V. Adaptive nested optimization scheme for multidimensional global 
+search / V. Gergel, V. Grishagin, A. Gergel // Journal of Global 
+Optimization. --- 2016. --- Vol. 66(1). --- P. 35--51.
 
-\bibitem{Sergeyev2017}
-Sergeyev, Ya.D. Deterministic global optimization: An introduction to the diagonal approach / Ya.D. Sergeyev, D.E. Kvasov --- New York: Springer, 2017. --- 136 p.
+\bibitem{Sergeyev2015}
+Sergeyev, Y.D.
+(2015) A deterministic global optimization using smooth diagonal auxiliary functions / Y.D.
+Sergeyev, D.E.
+Kvasov // Communications in Nonlinear Science and Numerical Simulation.
+--- 2015.
+--- Vol.
+21 (1-3).
+--- P.
+99--111.
 

### 2017-11-09 2018/ITNT/ITNT2018.tex
computations -> computing

@@
 
-The organization of the parallel computations with the use of the recursive 
+The organization of parallel computing with the use of the recursive 
 optimization scheme has been considered in details for the shared/distributed 

### 2017-11-10 2018/LaTeX/PAVT2018_MAM.tex
Added a folder remotely

--- 2017-10-25 2018/LION/LION2018.tex
@@
-\documentclass{llncs}
+%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% This is the LaTeX source for the instructions to authors using
+% the LaTeX document class 'llncs.cls' for contributions to
+% the Lecture Notes in Computer Sciences series.
+% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
+%
+% It may be used as a template for your own input - copy it
+% to a new file with a new name and use it as the basis
+% for your article.
+%
+% NB: the document class 'llncs' has its own and detailed documentation, see
+% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\usepackage{hyperref}
+\documentclass[runningheads,a4paper]{llncs}
+
+\usepackage{amssymb,amsmath}
+\setcounter{tocdepth}{3}
 \usepackage{graphicx}
-\usepackage{multirow}
-\usepackage[misc,geometry]{ifsym}
-
-\usepackage{amsmath}
 \usepackage{enumitem}
-%\usepackage{amsfonts}
-%\usepackage{amssymb}
-%\usepackage{epstopdf}
-%\usepackage{epsfig}
+\usepackage{url}
 
@@
 
-\mainmatter
+\mainmatter  % start of an individual contribution
 
-\title{Multipoint Approximation Method for Large Scale Design Optimization Problems Under Uncertainty}
-\author{Vassili Toropov \inst{1,3} \and Yury Korolev \inst{2} \and Victor Gergel \inst{3} \and Konstantin Barkalov \inst{3} %\Letter 
-\\
-\email{konstantin.barkalov@itmm.unn.ru}}
+% first the title is needed
+\title{Parallel Multipoint Approximation Method for Large Scale Optimization Problems 
+\thanks{This study was supported by the Russian Science Foundation, project No. 16-11-10150.}}
 
-\institute{
-Queen Mary University of London, London, UK
-\and
-University of L\"ubeck, Germany
-\and
-Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
+% a short form should be given in case it is too long for the running head
+\titlerunning{Parallel Multipoint Approximation Method}
+
+\author{ Victor P. Gergel%(ORCID 0000-0002-4013-2329)
+$^1$ \and
+Konstantin A. Barkalov%(ORCID 0000-0001-5273-2471)
+$^1$ \and
+Evgeny A. Kozinov%(ORCID )
+$^1$ \and
+Vassili V. Toropov%(ORCID 0000-0002-7017-5055)
+$^{1,2}$
+}
+%
+\authorrunning{V.P. Gergel \and K.A. Barkalov \and E.A.Kozinov \and V.V. Toropov}
+% (feature abused for this document to repeat the title also on left hand pages)
+
+% the affiliations are given next; don't give your e-mail address
+% unless you accept that it will be published
+\institute{$^1$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
+\email{\{victor.gergel, konstantin.barkalov, evgeny.kozinov\}@itmm.unn.ru}\\
+$^2$Queen Mary University of London, London, UK\\
+\email{v.v.toropov@qmul.ac.uk}\\
 }
 
+%
+% NB: a more complex sample for affiliations and the mapping to the
+% corresponding authors can be found in the file "llncs.dem''
+% (search for the string "\mainmatter" where a contribution starts).
+% "llncs.dem" accompanies the document class "llncs.cls".
+%
+
+%\toctitle{Lecture Notes in Computer Science}
+%\tocauthor{Authors' Instructions}
 \maketitle
@@
 \begin{abstract}
-The paper presents a new development in the Multipoint Approximation Method (MAM) that makes it capable of handling problems with uncertainty in design variables as well as in additional ‘environmental’ variables.
-The approach relies on approximations built in the combined space of design variables and environmental variables, and subsequent application of a risk measure and optimization with respect to the deterministic design variables, all within the iterative trust-region-based framework of MAM.
+TBD
 
-\keywords global optimization, design optimization, multidisciplinary optimization, uncertainty, multipoint approximation method.
-
+\keywords{design optimization, multidisciplinary optimization, multipoint approximation method, parallel computing.}
 \end{abstract}
@@
 \section{Introduction}
-\label{sec:intro}
-In industrial design optimization problems, random inaccuracies of the production process may lead to the designer’s inability to fully control the design variables.
-Therefore, there can be a discrepancy between the products \textit{as designed} and \textit{as produced} that results in actual performance being different from the performance expected at the design stage.
-Possible discrepancy between the design and the manufactured product becomes even more important if the latter violates some critical constraints, e.g.
-safety regulations.
-
-In addition, the system’s performance may depend on parameters of random nature that are beyond the designer’s control, e.g.
-the operation conditions.
-Therefore,  accounting for both of these types of uncertainties  is crucial for achieving robust and reliable performance.
-
-Optimization under uncertainty deals with responses $F(\pmb x,\pmb y)$ that depend on deterministic design variables $\pmb x$ and random ‘environmental’ variables $\pmb y$.
-Uncertainty in the design variables can also be modelled by introducing additional random variables, e.g.
-as additive or multiplicative noise.
-In this case, the responses $F(\pmb x,\pmb y)$ receive as input the sum $\pmb x+\pmb u$ (if the noise is additive), where $\pmb u$ is a random variable with a known probability distribution, yielding responses depending on $\pmb x,\; \pmb u$ and $\pmb y$.
-Randomness in $\pmb u$ and $\pmb y$ induces randomness in the responses $F(\pmb x+\pmb u,\pmb y)$.
-To convert them into deterministic functions, that can be optimized, a mapping $R$, called the \textit{risk measure} \cite{RockafellarUryasev2000} or \textit{robustness measure} \cite{SchillingsSchulz2015}, is applied to the responses.
-The result is then a deterministic function of the design variables: $\widetilde{F}(\pmb x)=R(F(\pmb x+\pmb u, \pmb y))$.
-The choice of a particular risk measure $R$ reflects the designer's attitude towards the risk of sub-optimal performance or constraint violation.
-For example, $R$ can be a combination of the mean and standard deviation of $F(\pmb x+ \pmb u,\pmb y)$ with respect to the joint probability distribution of $\pmb u$ and $\pmb y$ or a quantile of the distribution of $F(\pmb x+ \pmb u,\pmb y)$, i.e.
-such a value that the random variable of $F(\pmb x+\pmb u,\pmb y)$ exceeds it with a given, typically small, probability.
-
-Computations of the risk measures typically involve evaluation of high-dimensional integrals, which becomes a challenging task if the number of design variables and/or environmental variables is large.
-The problem becomes even more difficult if the original responses are computationally expensive, which is the case, for example, in computational fluid dynamics (CFD), where one function evaluation may take hours of even days.
-Therefore, direct computation of the risk measures using original responses becomes infeasible and therefore approximations of responses, also referred to as metamodels, need to be used.
+TBD
 
@@
 
-This paper presents a new development in the Multipoint Approximation Method that makes it capable of handling problems with uncertainty in design variables as well as in additional ``environmental'' variables.
-The approach relies on approximations built in the combined space of design variables and environmental variables, and subsequent application of a risk measure and optimization with respect to the deterministic design variables, all within the iterative trust-region-based framework of MAM.
+TBD
 
@@
 
-It is useful to start with a brief description of the deterministic version of MAM. A typical formulation of a constrained optimization problem that MAM works with is as follows:
+It is useful to start with a brief description of MAM. A typical formulation of a constrained optimization problem that MAM works with is as follows:
 \begin{equation}
@@
   \begin{array}{c}
-  \min\limits_{a_i \le x_i \le b_i}F_0(\pmb x) \\
-  s.t.\; F_j(\pmb x) \le 1,\; j=1,\dots ,M,
+  \min\limits_{a_i \le x_i \le b_i}F_0(x) \\
+  s.t.\; F_j(x) \le 1,\; j=1,\dots ,M,
   \end{array}
 \end{equation}
-where $\pmb x$ is the vector of design variables, $\pmb a$ and $\pmb b$ are the lower and upper bounds for the design variables, respectively, $F_0(\pmb x)$ is the objective function, and $F_j(\pmb x)$ are the constraints.
+where $x$ is the vector of design variables, $a$ and $b$ are the lower and upper bounds for the design variables, respectively, $F_0(x)$ is the objective function, and $F_j(x)$ are the constraints.
 The numbers of design variables and constraints are $n$ and $M$, respectively.
@@
 
-Algorithm 1 (deterministic MAM).
+Algorithm (MAM).
 \begin{enumerate}
-\item Initialization: choose a starting point $\pmb x^0$ and initial trust region $[\pmb a^0, \pmb b^0]$ such that $\pmb x^0 \in [\pmb a^0, \pmb b^0]$.
-\item At the $k$-th iteration the current approximation to the constrained minimum is $\pmb x^k$, the current trust region is $[\pmb a^k, \pmb b^k] \subset [\pmb a^0, \pmb b^0]$.
+\item Initialization: choose a starting point $ x^0$ and initial trust region $[ a^0,  b^0]$ such that $ x^0 \in [ a^0,  b^0]$.
+\item At the $k$-th iteration the current approximation to the constrained minimum is $ x^k$, the current trust region is $[ a^k,  b^k] \subset [ a^0,  b^0]$.
   \begin{enumerate}[label=(\alph*)]
-    \item Design of Experiments (DoE): a set of points $\pmb x_k^i \in [\pmb a^k, \pmb b^k]$ is chosen to be used for approximation building.
+    \item Design of Experiments (DoE): a set of points $ x_k^i \in [ a^k,  b^k]$ is chosen to be used for approximation building.
 Responses are evaluated at the DoE points and approximations are built using the obtained values.
@@
 
-    Denote the approximate objective function and constraints by $\widetilde{F}_0(\pmb x)$ and $\widetilde{F}_j(\pmb x)$, respectively.
+    Denote the approximate objective function and constraints by $\widetilde{F}^k_0(x)$ and $\widetilde{F}^k_j(x)$, respectively.
     \item The original optimization problem (\ref{eq:problem}) is replaced by the following problem:
@@
       \begin{array}{c}
-      \min\limits_{a_i^k \le x_i ^k\le b_i}\widetilde{F}_0(\pmb x) \\
-      s.t.\; \widetilde{F}_j(\pmb x) \le 1,\; j=1,\dots ,M,
+      \min\limits_{a_i^k \le x_i ^k\le b_i}\widetilde{F}^k_0(x) \\
+      s.t.\; \widetilde{F}^k_j( x) \le 1,\; j=1,\dots ,M,
       \end{array}
@@
     The approximate problem (\ref{eq:problem_approx}) is solved using Sequential Quadratic Programming (SQP) and the solution of this problem determines the center of the next trust region.
-    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region \cite{KeulenToropovMarkine1996}.
-    \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $\pmb x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations).
+    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $ x^k$, and on the size of the current trust region \cite{KeulenToropovMarkine1996}.
+    \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $ x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations).
 If the termination criterion is satisfied, the algorithm proceeds to step 3.
@@
   \end{enumerate}
-  \item Optimization terminates. The obtained approximation to the solution of the problem (\ref{eq:problem}) is $\pmb x^{k+1}$.
+  \item Optimization terminates. The obtained approximation to the solution of the problem (\ref{eq:problem}) is $ x^{k+1}$.
 \end{enumerate}
 
-\section{Optimization Under Uncertainty}
+The selection of the approximations $\widetilde{F}^k_j(x),\; j=0, \dots ,M,$ is such that their evaluation is inexpensive as compared to the evaluation of the original response functions $F_j(x)$.
+For example, intrinsically linear functions were successfully used for a variety of design optimization problems in the works \cite{ToropovFilatov1993,KeulenToropov1997}.
+The approximations are determined by means of the weighted least squares:
+\begin{equation}\label{eq:least_sq}
+\min \sum_{p=1}^P{w_{pj}\left[ F_j(x_p)- \widetilde{F}^k_j(x_p,a_j) \right]^2}.
+\end{equation}
+In (\ref{eq:least_sq}) minimization is carried out with respect to the tuning parameters $a_j$; $w_{pj}$ are the weight coefficients, and $P$ is the number of sampling points in Design of Experiments (DoE) which must not be less than the number of parameters in the vector $a_j$.
 
-The presence of uncertainties in the responses requires certain reformulations in the problem statement (\ref{eq:problem}).
-Informally, the optimization problem with uncertainty can be written as follows:
-\begin{equation}
-  \label{eq:problem_unc}
-  \begin{array}{c}
-  \min\limits_{a_i \le x_i \le b_i}F_0(\pmb x + \pmb u, \pmb y) \\
-  s.t.\; F_j(\pmb x + \pmb u, \pmb y) \le 1,\; j=1,\dots ,M,
-  \end{array}
+The weight coefficients $w_{pj}$ strongly influence the difference in the quality of the approximations in different regions of the design variable space.
+Since in realistic constrained optimization problems the optimum point usually belongs to the boundary of the feasible region, the approximation functions should be more accurate in such domain.
+Thus, the information at the points located near the boundary of the feasible region is to be treated with greater weights.
+In a similar manner a larger weight can be allocated to a design with a better objective function, see \cite{ToropovFilatov1993,KeulenToropov1997}.
+
+As optimization steps are carried out, a database with response function values becomes available.
+In order to achieve good quality approximations in the current trust region, an appropriate selection of DoE points must be made.
+In this work, DoE points in each trust region is generated randomly.
+Generally, points located far from the current trust region would not contribute to the improvement of the quality of the resulting approximations in the trust region.
+For this reason only points located in the neighborhood of the current trust region are taken into account.%, as depicted in Fig.
+1.
+A box in the space of design variables, which is approximately 1.5 to 1.8 times larger than the box representing the current trust region, was found by numerical experimentation to be a reasonable choice for the size of the neighborhood.
+
+In this work an approach is used that is based on the assembly of different approximate models $\{\varphi_l\}$ into one metamodel using the following form (note that the indices $j$ and $k$ are suppressed to simplify notation):
+\begin{equation}\label{eq:assembly}
+\widetilde{F}(x) = \sum_{l=1}^{NF}{b_l\varphi_l(x)},
 \end{equation}
-where the random variable $\pmb u$ represents (additive) noise in the design variables and $\pmb y$ is the environmental variable, representing input to the responses that cannot be influenced by the designer.
-This problem statement is not precise since the random responses $F_0 (\pmb x+\pmb u,\pmb y),\; F_j (\pmb x+\pmb u,\pmb y)$ cannot be directly optimized.
-Therefore, a risk measure $R$ needs to be applied to the objective and the constraints:
-\begin{equation}
-  \label{eq:problem_unc_r}
-  \begin{array}{c}
-  \min\limits_{a_i \le x_i \le b_i}R_{\pmb u,\pmb y}(F_0)(\pmb x) \\
-  s.t.\; R_{\pmb u,\pmb y}(F_j)(\pmb x) \le 1,\; j=1,\dots ,M,
-  \end{array}
-\end{equation}
-where the subscript in $R_{\pmb u,\pmb y}$ indicates that the variables $\pmb u$ and $\pmb y$ ‘collapse’ when the risk measure is applied and the result only depends on $\pmb x$.
-The choices of risk measure will be discussed in Section \ref{sec:risk}.
-Here the changes in the Algorithm 1 that are necessary to solve the problem (\ref{eq:problem_unc_r}) are discussed.
+where $NF$ is the number of regressors in the model pool $\{\varphi_l\}$ and $b_l$ are corresponding regression coefficients.
+The used procedure consists of two subsequent steps.
+In the first step, the parameters $a_l$ of individual functions (regressors) $\varphi_l$ in (\ref{eq:assembly}) are determined by solving a weighted least squares problem using a specified a DoE of $P$ points:
+%\begin{equation}\label{eq:least_sq}
+\[
+\min \sum_{p=1}^P{w_{p}\left[ F(x_p)- \varphi_l(x_p,a_l) \right]^2},
+\]
+%\end{equation}
+where minimization is carried out with respect to the tuning parameters $a_l$.
 
-First, it needs to be decided on whether to build metamodels for the risk measure $R_{\pmb u,\pmb y}(F_j)(\pmb x)$ or for the original responses $F_j (\pmb x+ \pmb u,\pmb y)$.
-The first option may seem attractive due to the simplicity of its implementation in Algorithm 1 but the deciding argument has to be that of the computational cost associated with the evaluation of either function.
-On the one hand, one evaluation of the function $R_{\pmb u,\pmb y}(F_j)(\pmb x)$ involves computing one or several integrals of the original response $F_j(\pmb x+ \pmb u,\pmb y)$ over the distribution of $\pmb u$ and $\pmb y$, which requires many evaluations of the original response.
-On the other hand, the dimensionality of the argument of $R_{\pmb u,\pmb y}(F_j)(\pmb x)$ is less than that of the argument of $F_j(\pmb x+ \pmb u,\pmb y)$, therefore, in principle, fewer function evaluations are needed to approximate the former than the latter.
-In practice, this may be a matter of a trade-off between the two abovementioned effects that depends on the proportion between the numbers of design variables $n$ and the environmental variables $n_e$ and on the complexity of the integration method and metamodeling method used.
-If the number if design variables is significantly larger than the number of environmental variables, which is often the case and which is assumed in this paper, building metamodels for the original responses $F_j(\pmb x+ \pmb u,\pmb y)$ is preferred.
+In the second step, based on the same DoE and keeping the obtained parameters $a_l$ fixed, a vector $b$ in (\ref{eq:assembly}) is estimated using the following formulation
+\[
+\min \sum_{p=1}^P{w_{p}\left[ F(x_p)- \widetilde{F}(x_p,b) \right]^2},
+\]
+that leads to solving a linear system of $NF$ equations with $NF$ unknowns $b_l$, where $NF$ is the number of regressors in the model pool $\{\varphi_l\}$.
 
-This choice requires some changes in the MAM algorithm, since metamodels are now built in a larger space than the one where optimization is performed in. The changes are summarized below.
-
-Algorithm 2 (MAM under uncertainty).
-\begin{enumerate}
-%  \item Initialization: choose a starting point $\pmb x_0 \in {\Bbb R}^n$. 
-  \item Initialization: choose a starting point $\pmb x_0 \in {\pmb R}^n$. 
-	Choose an initial trust region $[\pmb a^0,\pmb b^0]$ in the combined space of design variables and environmental variables. 
-	Ensure that the initial guess $\pmb x_0$ belongs to the projection of the initial trust region onto the subspace of design variables.
-The projection of the initial trust region onto the subspace of environmental variables must cover the region where the realizations of the random variable y are likely to be.
-The coordinates of the trust region corresponding to the environmental variables will not be changed in the course of optimization.
-Typically, the trust region along the coordinates corresponding to the $i$-th environmental variable is centered at the mean of the $i$-th environmental variable and has the width of six of its standard deviations.
-  \item At the $k$-th iteration the current approximation to the constrained minimum is $\pmb x^k$, the current trust region is $[\pmb a^k,\pmb b^k]$.
-  \begin{enumerate}[label=(\alph*)]
-    \item Design of Experiments (DoE): a set of points $(x_i^k,y_i^k)\in [\pmb a^k,\pmb b^k]$ is chosen that will be used to build approximations.
-Responses are evaluated at the DoE points and approximations are built using the obtained values.
-    Denote the approximate objective function and constraints by $\widetilde{F}_0(\pmb x, \pmb y)$ and $\widetilde{F}_j(\pmb x, \pmb y)$, respectively.
-    \item The original optimization problem is replaced by the following problem:
-    \begin{equation}
-      \label{eq:problem_u_tilde}
-      \begin{array}{c}
-      \min\limits_{a_i \le x_i \le b_i}R_{\pmb u,\pmb y}(\widetilde{F}_0)(\pmb x) \\
-      s.t.\; R_{\pmb u,\pmb y}(\widetilde{F}_j)(\pmb x) \le 1,\; j=1,\dots ,M,
-      \end{array}
-    \end{equation}
-    The risk measures are calculated using the metamodels obtained in the previous step.
-Integrals that arise during this process are calculated with respect to the joint distribution of $\pmb u$ and $\pmb y$.
-This distribution is assumed to be fixed.
-    The approximate problem (\ref{eq:problem_u_tilde}) is solved using Sequential Quadratic Programming (SQP) and the center of the next trust region $\pmb x^{k+1}$ is determined as the solution of this problem.
-The optimization runs in a subspace of the space where metamodels are built: metamodels are defined in the combined space of design variables and environmental variables, whilst optimization runs in the subspace of the design variables.
-    \item The trust region is updated.
-The size of the trust region along the dimensions corresponding to the environmental variables is unchanged.
-The size of the next trust region along the dimensions corresponding to the design variables is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region according to the trust region strategy explained in \cite{KeulenToropovMarkine1996}.
-    \item The termination criterion is checked. If the termination criterion is satisfied, the algorithm proceeds to step 3. Otherwise, it returns to step 2
-  \end{enumerate}
-  \item Optimization terminates. The obtained approximation to the optimum is $\pmb x^{k+1}$.
-\end{enumerate}
-
-Summarizing, the key changes introduced to account for uncertainties are (i) metamodels built in the combined space of design variables and environmental variables; (ii) optimization performed in a subspace of the space where metamodels are defined; (iii) additive noise in the design variables; (iv) application of risk measures to transform random responses into deterministic functions.
-
-\section{Risk Measures}
-\label{sec:risk}
-
-As it has been already said, a risk measure is a mapping that transforms a random response $F(\pmb x+ \pmb u,\pmb y)$  into a deterministic function that can be optimized.
-Next we provide some examples of risk measures \cite{RockafellarUryasev2000,SchillingsSchulz2015}.
-
-\begin{itemize}
-  \item $R(F)=\mu(F)$ where $\mu(F)$ is the mean of $F$.
-This choice corresponds to optimizing average performance of the system.
-In typical engineering applications, however, this is not sufficient and more conservative risk measures need to be chosen.
-  \item $R(F)=\sup(F)$.
-This is the most conservative case, in which one is concerned with the worst case performance of the system.
-The expression $\sup(F)$ stands for the least of all values that are greater than the random response $F$ with probability 1.
-If the support of the probability density function (PDF) of $F$ is not compact (i.e.
-$F$ can take arbitrary large values with non-zero probability) than $\sup(F)$ will be infinite.
-  \item $R(F)=\mu(F)+k\sigma(F)$, where $\sigma(F)$ is the standard deviation of $F$ and $k>0$ is a constant.
-This risk measure is sometimes referred to as the safety margin.
-Optimization using this risk measure is referred to as robust design optimization \cite{KochYangGu2004,Yao2011,ZhangZhuChenArendt2013} and reflects the paradigms of a ‘$3\sigma$-design’, a ‘$6\sigma$-design’, etc.
-  \item $R(F)=q_F(\alpha)$, where $q_F(\alpha)$ is the value of the quantile function, or simply the quantile, corresponding to a given probability $\alpha\in (0,1)$.
-This choice corresponds to the paradigm of reliability-based design optimization  \cite{AouesChateauneuf2010,Yao2011,YounChoi2004}.
-In this approach, one wants to keep the probability of the system failing (i.e.
-violating the constraint or underperforming, depending on the context) below a given threshold.
-The value $1-\alpha$ is the desired probability of failure.
-Although rather intuitive, this choice has known computational issues which are discussed by Rockafellar and Royset \cite{RockafellarRoyset2010}.
-\end{itemize}
-
-For this study, we chose as risk measure the safety margin $R(F)=\mu(F)+k\sigma(F)$, aiming at covering both robust design optimization and reliability design optimization.
-Let us briefly discuss its computation.
-
-To compute the safety margin risk measure, one needs to compute two integrals:
-\begin{displaymath}
-  \begin{array}{c}
-    \mu (F)=E[F]=\int F(\pmb x + \pmb u, \pmb y)p df(\pmb u, \pmb y)d\pmb u d\pmb y \\
-    \sigma^2(F)=E[(F-E[F])^2]=\\\int(F(\pmb x + \pmb u, \pmb y)-E[F])^2 p df(\pmb u, \pmb y)d\pmb u d\pmb y
-  \end{array}
-\end{displaymath}
-
-The dimension of these integrals is $n+n_e$.
-To compute them efficiently, methods of high-dimensional integration must be employed.
-Here, quasi-Monte-Carlo integration \cite{Caflisch1998} based on Niederreiter sequences \cite{Niederreiter1992} is used.
+\section{Parallel Algorithm}
+\label{sec:par_alg}
 
@@
   \begin{array}{c}
-    \min\limits_{\pmb b, \pmb h}V(\pmb b, \pmb h) \\
+    \min\limits_{ b,  h}V( b,  h) \\
     s.t.\;1.0\le b_i \le 10.0, \\
@@
 
-Now let us add uncertainty to the problem, adding Gaussian noise with zero mean and variance 0.1 to each design variable.
-Using the $\mu(F) + k\sigma(F)$ risk measure, we obtain the following problem statement:
-\begin{displaymath}
-  \begin{array}{c}
-    \min\limits_{\pmb b, \pmb h}\mu(V(\pmb b, \pmb h) + k\sigma V(\pmb b, \pmb h)) \\
-    s.t.\;1.0\le b_i \le 10.0, \\
-    5.0 \le h_i \le 100.0, \\
-    \mu(y_N) + k\sigma(y_N)\le 2.5, \\
-    \mu(\sigma_i)+k\sigma(\sigma_i)\le \bar{\sigma}=14000, \\
-    \mu(\frac{h_i}{b_i})+k\sigma(\frac{h_i}{b_i})\le 20.
-  \end{array}
-\end{displaymath}
-
-The notation $\sigma$ was used to denote the variance of a response, but it also stands for the maximum bending stresses $\sigma_i$, resulting in somewhat confusing notation $\sigma(\sigma_i )$.
-Hopefully, this explanation eliminates the confusion.
-Optimizing $\mu(F) + k\sigma(F)$ with $k = 3.0$ using SQP on original functions, we obtain the value of the deterministic objective (volume of the structure) at the robust optimum $V = 67598.049\; cm^3$.
-To compute $\mu(F)$ and $\sigma(F)$ at each iteration, 1000 points were used.
-Multiplied by the number of evaluations of the robust objective $\mu(F) + k\sigma(F)$ (approximately 8000), we get a rough estimate of $8\cdot 10^6$ of the total number of points that SQP used to find this solution.
-
-Solving the same problem using the new version of MAM suitable for problems with uncertainty, we obtain the value of the deterministic objective (volume of the structure) at the robust optimum $V = 67799.335\; cm^3$.
-To obtain this result, MAM again used 150 points per trust region to build metamodels, which resulted in a total of 1500 function evaluations, which is orders of magnitude smaller than the number of points required if no metamodels are used.
-
-The $\mu(F ) + k\sigma(F)$ risk measure uses quantities that are estimated in the middle of the distribution --- the mean and the standard deviation.
-On the one hand, this has the advantage that samples drawn directly from the distribution of the random variables $\pmb u$ and $\pmb y$ are well suited to approximately compute this risk measure.
-On the other hand, probabilistic interpretation of the obtained results in this case is only possible under the normality assumption for the distribution of $F$; if the actual distribution of $F$ has heavier tails than the normal one, the actual probability of failure will be higher than the one  the system was designed for.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-Recent developments in the Multipoint Approximation Method made it capable of solving large scale problems with uncertainty in parameters modelled as additional (``environmental'') random variables and uncertainty in the design variables modelled as additive noise.
-The approach taken combines metamodels built in the combined space of design variables and environmental variables, application of risk measures that map random responses to deterministic functions, and optimization in a subspace of the space where metamodels are defined, all within the iterative and trust-region-based framework of MAM.
-Performance is demonstrated on a benchmark example of structural optimization known as the scalable cantilevered beam.
-
-\section*{Acknowledgment}
-The authors are grateful for the support provided by the Russian Science Foundation, project No. 16-11-10150.
+\section{Conclusions}
 
@@
 
+%\bibitem{Strongin2000}
+%Strongin, R.G., Sergeyev, Y.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000); DOI: 10.1007/978-1-4615-4677-1
+%
+%\bibitem{Barkalov2002}
+%Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)
+%
+%\bibitem{Gergel2005}
+%Gergel, V.P., Strongin, R.G.: Parallel computing for globally optimal decision making on cluster systems. Future Gener. Comput. Syst. 21(5), 673--678 (2005); DOI: 10.1016/j.future.2004.05.007
+%
+%\bibitem{Strongin2013}
+%Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013);  DOI: 10.1007/978-1-4614-8042-6
+%
+%\bibitem{Barkalov2010}
+%Barkalov, K., Ryabov, V., Sidorov, S.: Parallel Scalable Algorithms with Mixed Local-Global Strategy for Global Optimization Problems.
+In: Hsu, Ch., Malyshkin, V.
+(Eds.) MTPP 2010, LNCS, vol.
+6083, pp.
+232--240.
+Springer, Heidelberg (2010); DOI: 10.1007/978-3-642-14822-4$\_$26
+%
+%\bibitem{Barkalov2015}
+%Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi Coprocessor for Solving Global Optimization Problems.
+In: Malyshkin, V.
+(Ed.) PaCT 2015, LNCS, vol.
+9251, pp.
+307--318.
+Springer, Heidelberg (2015); DOI: 10.1007/978-3-319-21909-7$\_$31
+%
+%\bibitem{Barkalov2016}
+%Barkalov, K., Gergel, V.: Parallel Global Optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016); DOI: 10.1007/s10898-016-0411-y
+%
+%\bibitem{Gergel2016}
+%Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster.
+In: Simos T.E.
+(Ed.) ICNAAM 2015, AIP Conference Proceedings, vol.
+1738, art.
+no.
+400006 (2016); DOI: 10.1063/1.4952194
+%
+%\bibitem{Gergel2015}
+%Gergel, V., Sidorov, S.: A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems.
+In: Malyshkin, V.
+(Ed.) PaCT 2015, LNCS, vol.
+9251, pp.
+505--515.
+Springer, Heidelberg (2015); DOI: 10.1007/978-3-319-21909-7$\_$49
+%
+
+%\end{thebibliography}
+
 \end{document}
-______________________________________________________________________

### 2017-11-13 2018/PAVT_Toropov/PAVT2018_MAM.tex
Text from Toropov's papers

--- 2017-11-10 2018/LaTeX/PAVT2018_MAM.tex
@@
 $^1$ \and
-Evgeny A. Kozinov%(ORCID )
+Evgeny A. Kozinov%(ORCID 0000-0001-6776-0096)
 $^1$ \and
@@
 Generally, points located far from the current trust region would not contribute to the improvement of the quality of the resulting approximations in the trust region.
-For this reason only points located in the neighborhood of the current trust region are taken into account.%, as depicted in Fig.
-1.
+For this reason only points located in the neighborhood of the current trust region are taken into account, as depicted in Fig.~\ref{fig:trustregion}.
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.5\textwidth]{trustregion.png}
+    \caption{Current trust region (smaller box) and its extension (larger box): points outside the larger box are not used for building the approximate functions}
+    \label{fig:trustregion}
+\end{figure}
 A box in the space of design variables, which is approximately 1.5 to 1.8 times larger than the box representing the current trust region, was found by numerical experimentation to be a reasonable choice for the size of the neighborhood.
@@
 that leads to solving a linear system of $NF$ equations with $NF$ unknowns $b_l$, where $NF$ is the number of regressors in the model pool $\{\varphi_l\}$.
+
+The selection of the regressors $\{\varphi_l\}$ is based on the number of sampling points currently located in the trust region.
+In the mid-range approximation framework, inexpensive approximate models for objective and constraint functions are built using minimum required number of sampling points.
+The simplest case is a linear function of the tuning parameters $a$:
+\[
+\varphi(x) = a_0 + \sum_{i=1}^N{a_i x_i}.
+\]
+
+This structure can be extended to an \textit{intrinsically linear} function.
+Such functions are nonlinear, but they can be led to linear ones by simple transformations.
+The most useful function among them is the multiplicative function
+\[
+\varphi(x) = a_0 \prod_{i=1}^N{x{_i}^{a_i}}.
+\]
+
+Intrinsically linear functions have been successfully used for a variety of design optimization problems. 
+The advantage of these approximation functions is that a relatively small number $N+1$ ($N$ is number of design variables) of tuning parameters $a_i$ is to be determined, and the corresponding least squares problem is solved easily.
+This is the most important feature of such approximations as it allows applying them to large scale optimization problems.
+
+Other intrinsically linear functions may be considered in the model pool, e.g.
+\[
+\varphi(x) = a_0 + \sum_{i=1}^N{a_i/x_i},
+\]
+\[
+\varphi(x) = a_0 + \sum_{i=1}^N{a_i x_i^2},
+\]
+\[
+\varphi(x) = a_0 + \sum_{i=1}^N{a_i/x_i^2},
+\]
+\[
+\varphi(x) = a_0 + \sum_{i=1}^N{a_i x_i^3},
+\]
+\[
+\varphi(x) = a_0 + \sum_{i=1}^N{a_i/x_i^3}.
+\]
+
+As more points are added to the database the approximations may be switched to higher quality models, e.g. a rational model
+\begin{equation}\label{eq:rational}
+\varphi(x) = \frac{a_1 + a_2x_1 + a_3x_2+...+a_{n+1}x_n}{1 + a_{n+2}x_1 + a_{n+3}x_2+...+a_{2n+1}x_n},
+\end{equation}
+The coefficients in (\ref{eq:rational}) are determined using a least squares approach which reduces to a nonlinear optimization problem with a constraint on the sign of the denominator (positive or negative).
+The latter is necessary in order to prevent the denominator from crossing the zero axis within a specified trust region.
+One may note that this formulation may yield the objective function with many local minima.
+Currently this problem is resolved using optimization restarts from a specified number of initial guesses randomly generated in a trust region.
+
+Tests results demonstrated that, although the above functions may describe the global behavior rather poorly, such approximations proved to be efficient in the mid-range approximation framework of MAM.
 
@@
 With $N=50$ segments (corresponding to 100 design variables) the SQP solution of this problem is $V = 63704.598\; cm^3$.
+The optimal values of design variables are given below (the first 50 entries are the widths and the last 50 entries are the heights of the segments):
+\setcounter{MaxMatrixCols}{11}
+\[
+\begin{matrix}%[pos]{spalten}
+			b=& 3.246& 3.224& 3.202& 3.179& 3.156& 3.133& 3.109& 3.085& 3.061& 3.036 \\  
+			  & 3.011& 2.985& 2.959& 2.933& 2.905& 2.878& 2.850& 2.821& 2.792& 2.762 \\
+			  & 2.731& 2.700& 2.668& 2.635& 2.602& 2.567& 2.532& 2.495& 2.458& 2.419 \\  
+			  & 2.379& 2.338& 2.295& 2.250& 2.204& 2.156& 2.105& 2.052& 1.996& 1.936 \\  
+			  & 1.873& 1.805& 1.732& 1.651& 1.562& 1.462& 1.345& 1.196& 1.023& 1.000 \\ 	
+\end{matrix}
+\]
+\[
+\begin{matrix}
+	h=& 64.919&64.480&64.033&63.581&63.122&62.656&62.184&61.703&61.216&60.720  \\
+		& 60.216&59.703&59.182&58.651&58.110&57.558&56.996&56.423&55.837&55.239  \\
+		& 54.628&54.003&53.363&52.707&52.035&51.344&50.635&49.906&49.155&48.381  \\
+		& 47.582&46.754&45.897&45.008&44.082&43.116&42.105&41.041&39.919&38.729  \\
+		& 37.462&36.102&34.630&33.020&31.240&29.241&26.910&23.919&20.465&14.639  \\
+\end{matrix}
+\]
+
 MAM obtained the solution $V = 63935.360\; cm^3$, using 1500 function evaluations (as compared to almost 8000 evaluations used by SQP).
 The number of points in a trust region used to build approximations was 150.
+The optimal values of design variables obtained by MAM are given below:
+\[
+\begin{matrix}
+	b=&	3.238&3.204&3.202&3.170&3.176&3.128&3.108&3.089&3.057&3.003 \\  
+		& 3.027&2.986&2.959&2.952&2.885&2.855&2.864&2.847&2.803&2.762 \\  
+		& 2.737&2.722&2.630&2.645&2.590&2.558&2.547&2.480&2.693&2.391 \\ 
+		& 2.368&2.310&2.307&2.227&2.176&2.149&2.106&2.016&2.007&1.925 \\  
+		& 1.864&1.843&1.758&1.635&1.582&1.934&1.332&1.173&1.026&2.419 \\ 
+\end{matrix}
+\]
+\[
+\begin{matrix}
+	h=& 64.764&64.083&64.036&63.392&63.518&62.566&62.153&61.772&61.149& 60.054  \\
+		& 60.534&59.728&59.182&59.033&57.695&57.105&57.274&56.943&56.057&55.233  \\
+		& 54.747&54.433&52.590&52.902&51.802&51.150&50.943&49.597&51.567&47.811  \\
+		& 47.373&46.190&46.140&44.543&43.525&42.991&42.119&40.308&40.141&38.506  \\
+		& 37.268&36.839&35.153&32.702&31.615&28.304&26.642&23.447&20.536& 9.417
+
+\end{matrix}
+\]
+
 The solution obtained by MAM is very close to the reference solution obtained by SQP, except for the last design variable (the height of the last segment), which indicates that the problem is insensitive to this variable near the optimum, making it hard for the metamodels to capture this dependence.

### 2017-11-14 2018/ITNT/ITNT2018.tex
version for submission

@@
 
-\address{Lobachevsky State University of Nizhni Novgorod, Gagarin avenue 23, 
-Nizhni Novgorod, Russia, 603950}
+\address{National Research Lobachevsky State University of Nizhny Novgorod, Gagarin avenue 23, 
+Nizhny Novgorod, Russia, 603950}
 
@@
 therefore, for solving such problems, the methods are required, which 
-generate an essentially nonuniform grid in the search domain, denser in the 
-nearness of the global minimum and more sparse far away from this one. In the 
-present paper, the results of investigation of the scaling of the parallel 
+generate a nonuniform grid in the search domain, denser near the 
+global minimizer and more sparse far away from this one. In the 
+present paper, the results of scalability investigation of the parallel 
 index global optimization algorithm developed in Lobachevsky State 
-University of Nizhni Novgorod \cite{Strongin2000,Strongin2013} and implemented in the Globalizer 
-system [NACO] are presented.
+University of Nizhny Novgorod \cite{Strongin2000,Strongin2013} and implemented in the Globalizer 
+system \cite{Globalizer} are presented.
 
@@
 
-In the algorithms developed in the present paper the objective function and constraints are assumed to be Lipschitzian that is typical for many other approaches to the development of the parallel optimization algorithms (see, for example, \cite{Jones,Zilinskas,Evtushenko}).
+In the algorithms considered in the present paper the objective function and constraints are assumed to be Lipschitzian that is typical for many other approaches to the development of the parallel optimization algorithms (see, for example, \cite{Jones,Zilinskas,Evtushenko}).
 This assumption is natural for many applied problems since the relative variations of the functions generally cannot exceed a certain threshold determined by the bounded energy of changes occurring in the system under study.
@@
 of the functions connected with each other. Then, the solving of the 
-``difficult'' part of the problem (for which the execution of the computation-
-costly algorithms is required to conduct the trials) can be performed on the 
-CPU using an efficient index global optimization algorithm. The ``easy'' part 
-of the problem (the execution of which doesn’t require the complex algorithms 
-and can be transferred to an accelerator easily) can be solved on a GPU using 
+``difficult'' part of the problem (for which the execution of time-consuming algorithms is required to conduct the trials) can be performed on CPU using an efficient index global optimization algorithm.
+The ``easy'' part
+of the problem (the execution of which doesn't require the complex algorithms 
+and can be transferred to an accelerator easily) can be solved on GPU using 
 the uniform grid technique. Obviously, in this case a major portion of the trials 
-will be performed on the GPUs, and the minor one -- on the CPU. However, 
+will be carried out on GPU, and the minor one will be carried out on CPU. However, 
 because of the difference in the difficulties of the subproblems solved on 
@@
 system for solving the global optimization problem developed in Lobachevsky 
-State University of Nizhni Novgorod.
+State University of Nizhny Novgorod.
 
@@
 \end{equation}
-unambiguously mapping the interval of real axis $[0,1]$ onto a $N$-
-dimensional cube is the first of the dimensionality reduction methods 
-considered.  To implement this method of dimensionality reduction a 
+unambiguously mapping the interval of real axis $[0,1]$ onto a $N$-dimensional cube is the first of the dimensionality reduction methods considered.
+To implement this method of dimensionality reduction a
 numerically constructed curve (\textit{evolvent}) is used. The evolvent is 
@@
 where the $i$-th block variable $u_i$ is a vector of $N_i$ components of 
-vector $y$, taken serially i.e. $u_1=(y_1,y_2,...,y_{N_1})$, 
+vector $y$, taken serially, i.e. $u_1=(y_1,y_2,...,y_{N_1})$, 
 $u_2=(y_{N_1+1},y_{N_1+2},...,y_{N_1+N_2})$,..., $u_M=(y_{N-N_M+1},y_{N-
@@
 memory as well as for the accelerators in \cite{Sysoyev,BarkalovLebedev}. However, in these works the 
-problems were considered, in which the time of the trial execution didn’t 
+problems were considered, in which the time of the trial execution didn't 
 depend on the trial point. Here, the problem is considered assuming that in 
 the function $\varphi(y_1,...,y_N)$ it is possible to select more difficult 
-part $f(y_1,...,y_N)$ (depending on all problem parameters) and a simpler 
-part $g(y_s,...,y_N )$ (depending on a part of the parameters only), for 
+part $f(y_1,...,y_s)$ (depending on a part of the parameters only) and a simpler 
+part $g(y_1,...,y_N)$ (depending on all problem parameters), for 
 example, $\varphi(y)= f(y)g(y)$.
@@
 computations related to the numerical simulation, which can be performed on the 
-CPUs only. The simple part doesn’t imply the complex computations and can be 
+CPUs only. The simple part doesn't imply the complex computations and can be 
 computed on an accelerator, for example, GPU. For solving a problem with such 
 a structure, one can apply the parallel recursive optimization scheme with 
-the use the CPUs at the upper nesting level and the GPUs at the lower one.
+the use CPU at the upper nesting level and GPU at the lower one.
 
@@
 implemented in the Globzlizer solver developed in Lobachevsky State 
-University of Nizhni Novgorod. The global search methods and various 
+University of Nizhny Novgorod \cite{Globalizer}. The global search methods and various 
 dimensionality reduction schemes make the algorithmic base for the 
@@
 (b). At the same time, computing the values of the function $f(y_1,y_2)$ is 
-more computation-costly than of $g(y_3,y_4)$.
+more time-consuming than of $g(y_3,y_4)$.
 
@@
 graphic accelerators. Thus, at $p=32$ total 96 GPU accelerators were employed 
-each including 2688 CUDA cores. In table 2, the execution time of the 
-algorithm is presented and in table 3 -- the speedup with respect to the run 
-on a single node.
+each including 2688 CUDA cores. In tables 2 and 3 the average time and the speedup with respect to the run 
+on a single node are presented.
 
@@
 Computational experiments were carried out on a high-performance cluster of 
-Lobachevsky State University of Nizhni Novgorod. The cluster node included 
+Lobachevsky State University of Nizhny Novgorod. The cluster node included 
 two Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM, and three NVIDIA 
 Kepler K20Õ GPUs (2688 CUDA cores, 6 Gb GDDR5). 
-The results of experiments demonstrate that the paralle index algorithm combined with the nested optimization scheme provides a good speedup on the considered problem class.
+The results of experiments demonstrate that the parallel index algorithm combined with the nested optimization scheme provides a good speedup on the considered problem class.
 
@@
 \section*{Acknowledgments}
-This study was supported by the Russian Science Foundation, project No 16-11-
-10150.
+This study was supported by the Russian Science Foundation, project No 16-11-10150.
 
@@
 of the Moscow State University, 2013. --- 280 p. --- (in Russian).
+
+\bibitem{Globalizer}
+Sysoyev, A.V.
+Globalizer -- A parallel software system for solving global optimization problems / A.V.
+Sysoyev, K.A.
+Barkalov, V.V.
+Sovrasov, I.G.
+Lebedev, V.P.
+Gergel // Lecture Notes in Computer Science.
+--- 2017.
+--- Vol.
+10421.
+--- P.
+492--499.
 

### 2017-11-15 2018/PAVT_Toropov/PAVT2018_MAM.tex
Text in Russian

@@
 \documentclass[runningheads,a4paper]{llncs}
-
 \usepackage{amssymb,amsmath}
@@
 \usepackage{url}
+
+\usepackage[T2A]{fontenc}
+\usepackage[utf8x]{inputenc}
+\usepackage[english, russian]{babel}
 
@@
 \section{Introduction}
-TBD
+
+\Russian
+В статье рассматривается the multipoint approximation method (MAM) и его применение к large scale optimization problems.
 
@@
 
-TBD
+\Russian
+This paper presents a Parallel Multipoint Approximation Method that makes it capable of handling problems with the number of design variables in the order of thousands.
+Параллельный вариант алгоритма был разработан с целью минимизировать время, затрачиваемое на работу непосредственно оптимизационной части алгоритма.
+При этом предполагается, что процесс вычисления значений целевой функции уже распараллелен.
 
@@
 
-\section{Parallel Algorithm}
-\label{sec:par_alg}
+\section{Parallel Multipoint Approximation Method}\label{sec:par_alg}
+
+\Russian
+Рассмотрим возможные способы распараллеливания, которые можно применить для рассматриваемой задачи.
+Во-первых, можно распараллеливать вычисление функций, описывающих оптимизируемый объект.
+Данный путь является как очевидным, так и необходимым, т.к.
+в industrial design optimization problems расчет даже одного значения функции может занимать несколько часов.
+Однако данный способ является специфичным для каждой конкретной задачи.
+Вопросы распараллеливания тут решаются на уровне пакетов прикладных программ, в которых выполняется инженерное моделирование (например, Ansys, OpenFOAM, и т.п.).
+
+Во-вторых, можно скорректировать алгоритм с целью параллельного вычисления нескольких значений целевой функции и ограничений в разных точках области поиска.
+В соответствии с правилами MAM на каждой итерации в текущем trust region формируется P sampling points in design of experiments.
+Значения функции в них можно вычислять параллельно на разных процессорах (узлах).
+Сказанное соответствует распараллеливаню шага 2а алгоритма.
+The number of sampling points generated in each iteration may be set equal to P = k×NP number, where NP is a number of available processors (or nodes) and k ≥ 1.
+In terms of the time, the latter will be equivalent to NP function evaluations per step.
+Данный способ был успешно реализован в работе [35] и показал хорошую эффективность, т.к.
+при числе design variables порядка ста основное время затрачивается на function evaluations, работа самого MAM (в последовательном режиме) вносит незначительные накладные расходы.
+
+Однако когда число переменных становится порядка тысячи, работа последовательной части MAM начинает оказывать ощутимое влияние на общее время решения задачи (в предположении, что время вычисления значений целевой функции не меняется).
+Так, для рассмотренной в Section 4 задачи, время проведения одной итерации метода увеличилось в более чем в 2000 раз (с 4 до 9000 секунд ) при увеличении числа переменных в задаче с 100 до 1000.
+
+Поэтому в рамках проводимого исследования был реализован третий подход к распараллеливанию алгоритма.
+Были распараллелены вычислительные правила МАМ, обеспечивающие построение аппроксимации, решение аппроксимированной задачи и выбор следующего trust region (шаги 2b, 2c и 2d алгоритма).
+
+С целью выявления наиболее трудоемких мест алгоритма нами был использован Intel VTune Amplifier XE.
+Проведенный анализ показал, что наиболее трудоемкими операциями, использованными при реализации MAM, являются матричное умножение и решение СЛАУ при построении аппроксимации, а также решение аппроксимированной задачи методом SQP.
+
+Матричное умножение и решение СЛАУ являются стандартными операциями, реализованными во многих высокопроизводительных библиотеках.
+Здесь нами были использованы соответствующие параллельные методы из библиотеки Intel MKL.
+Метод SQP был распараллелен самостоятельно с использованием технологии OpenMP.
 
@@
 
+\Russian
+Далее сравним время работы последовательной и параллельной версий программы при решении large scale problems.
+Размерность N будем варьировать от 100 до 1000 design variables, что соответствует 50 – 500 сегментам cantilevered beam.
+The number of points in a trust region used to build approximations was 1.5 N  Таблица 1 отражает время работы параллельного алгоритма в зависимости от числа переменных.
+В таблице 2 приведено число итераций MAM, а также число вычислений значений функции, потребовавшееся для решения задачи.
+Для сравнения приведено также число вычислений значений функции, которые бы потребовались для решения исходной (неаппроксимированной) задачи методом SQP.
+Во всех проведенных экспериментах значения целевой функции в последовательной и параллельной версиях алгоритма совпадают (с точностью до вычислительной погрешности), и не сильно отличаются от решения, полученного SQP.
+При этом наглядно демонстрируется сокращение числа function evaluations, требующееся для решения задачи с использованием MAM по сравнению с использованием SQP.
+
+%Таблица
+
+Computational experiments were carried out on a high-performance cluster of Lobachevsky State University of Nizhny Novgorod.
+The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.
+The CPU had 8 cores, i.e.
+a total of 16 physical cores were available in the node.
+MS Visual Studio 15 and Intel Fortran Compiler were used to implement the algorithm.
+
 \section{Conclusions}
+
+Recent developments in the Multipoint Approximation Method (MAM) made it capable of solving large scale industrial optimization problems.
+The approach uses metamodels built in the space of design variables.
+MAM solves initial problem by using approximations of the objective function and constraints.
+\Russian
+В рамках данного исследования разработана параллельная версия МАМ, ориентированная на сокращение времени работы оптимизационного алгоритма (в предположении, что вычисление значений the objective function and constraints уже распараллелены).
+Проведенные эксперименты показывают приемлемое ускорение при решении задач большого размера при использовании 16 ядер.
+
+Performance is demonstrated on a benchmark example of structural optimization known as the scalable cantilevered beam.
 

### 2017-11-16 2018/PAVT_Toropov/PAVT2018_MAM.tex
Text in Russian

@@
 \usepackage[utf8x]{inputenc}
-\usepackage[english, russian]{babel}
+\usepackage[russian,english]{babel}
 
@@
 \begin{abstract}
-TBD
+The paper presents a new development in the Multipoint Approximation Method (MAM) that makes it capable of handling large scale problems.
+The approach relies on approximations built in the space of design variables within the iterative trust-region-based framework of MAM.
+\Russian
+С целью решения за приемлемое время задач большой размерности разработан параллельный варинат of the Multipoint Approximation Method (PMAM).
+При этом предполагается, что значения целевой функции и ограничений вычисляются с использованием distributed memory (на нескольких узлах кластера), а оптимизационный модуль работает на одном узле с использованием shared memory.
+Проведены вычислительные эксперименты on a benchmark example of structural optimization.
 
@@
 This paper presents a Parallel Multipoint Approximation Method that makes it capable of handling problems with the number of design variables in the order of thousands.
-Параллельный вариант алгоритма был разработан с целью минимизировать время, затрачиваемое на работу непосредственно оптимизационной части алгоритма.
-При этом предполагается, что процесс вычисления значений целевой функции уже распараллелен.
+Параллельный вариант алгоритма (с использованием shared memory) был разработан с целью минимизировать время, затрачиваемое на работу части алгоритма, связанной с построением аппроксимации и решением аппроксимированной задачи и не связанной с вычислением значений целевой функции и ограничений.
+Предполагается, что процесс вычисления значений целевой функции и ограничений (с использованием distributed memory) уже распараллелен.
 
@@
 Значения функции в них можно вычислять параллельно на разных процессорах (узлах).
-Сказанное соответствует распараллеливаню шага 2а алгоритма.
+Сказанное соответствует распараллеливаню шага 2а алгоритма с использованием distributed memory.
 The number of sampling points generated in each iteration may be set equal to P = k×NP number, where NP is a number of available processors (or nodes) and k ≥ 1.
@@
 Однако когда число переменных становится порядка тысячи, работа последовательной части MAM начинает оказывать ощутимое влияние на общее время решения задачи (в предположении, что время вычисления значений целевой функции не меняется).
-Так, для рассмотренной в Section 4 задачи, время проведения одной итерации метода увеличилось в более чем в 2000 раз (с 4 до 9000 секунд ) при увеличении числа переменных в задаче с 100 до 1000.
+Так, для рассмотренной в Section 4 задачи, время проведения одной итерации метода увеличилось в более чем в 2000 раз (с 4 до 9000 секунд ) при увеличении числа переменных со 100 до 1000.
 
-Поэтому в рамках проводимого исследования был реализован третий подход к распараллеливанию алгоритма.
+Поэтому в рамках проведенного исследования был реализован третий подход к распараллеливанию алгоритма.
 Были распараллелены вычислительные правила МАМ, обеспечивающие построение аппроксимации, решение аппроксимированной задачи и выбор следующего trust region (шаги 2b, 2c и 2d алгоритма).
 
-С целью выявления наиболее трудоемких мест алгоритма нами был использован Intel VTune Amplifier XE.
+С целью выявления наиболее трудоемких мест разработанной ранее последовательной программы нами был использован Intel VTune Amplifier XE.
 Проведенный анализ показал, что наиболее трудоемкими операциями, использованными при реализации MAM, являются матричное умножение и решение СЛАУ при построении аппроксимации, а также решение аппроксимированной задачи методом SQP.
@@
 Здесь нами были использованы соответствующие параллельные методы из библиотеки Intel MKL.
-Метод SQP был распараллелен самостоятельно с использованием технологии OpenMP.
+Метод SQP был распараллелен нами самостоятельно с использованием технологии OpenMP.
 
@@
 \Russian
-Далее сравним время работы последовательной и параллельной версий программы при решении large scale problems.
-Размерность N будем варьировать от 100 до 1000 design variables, что соответствует 50 – 500 сегментам cantilevered beam.
-The number of points in a trust region used to build approximations was 1.5 N  Таблица 1 отражает время работы параллельного алгоритма в зависимости от числа переменных.
-В таблице 2 приведено число итераций MAM, а также число вычислений значений функции, потребовавшееся для решения задачи.
-Для сравнения приведено также число вычислений значений функции, которые бы потребовались для решения исходной (неаппроксимированной) задачи методом SQP.
-Во всех проведенных экспериментах значения целевой функции в последовательной и параллельной версиях алгоритма совпадают (с точностью до вычислительной погрешности), и не сильно отличаются от решения, полученного SQP.
+Далее сравним время работы последовательного и параллельного алгоритмов при решении large scale problems. 
+Размерность решаемой задачи N варьировалась от 100 до 1000 design variables, что соответствует 50 – 500 сегментам cantilevered beam.
+The number of points in a trust region used to build approximations was 1.5 N.
+Оба алгоритма запускались на одном узле кластера (характеристики узла приведены ниже), при этом параллельный алгоритм использовал все доступные 16 процессорных ядер.
+Так как в решаемой тестовой задаче время вычисления значений целевой функции и ограничений является незначительным, то они вычислялись на том же узле.
+Table 1 отражает время работы (в секундах) последовательного и параллельного алгоритмов в зависимости от числа переменных.
+В table 2 приведено число итераций MAM $I_{MAM}$, а также число function evaluations $F_{MAM}$, потребовавшееся для решения задачи.
+Для сравнения приведено также число вычислений значений функции, которые бы потребовались для решения исходной (неаппроксимированной) задачи методом SQP $F_{SQP}$.
+Во всех проведенных экспериментах значения целевой функции в последовательной и параллельной версиях алгоритма совпадают (с точностью до вычислительной погрешности), и незначительно отличаются от решения, полученного SQP.
 При этом наглядно демонстрируется сокращение числа function evaluations, требующееся для решения задачи с использованием MAM по сравнению с использованием SQP.
@@
 %Таблица
+\begin{table}
+	\caption{Time and speedup}
+	\label{tab:1}
+	\center
+	\begin{tabular}{cccc}
+		\hline\noalign{\smallskip}
+		$N$ & MAM time & PMAM time & Speedup \\
+		\hline\noalign{\smallskip}
+		100 & 48 &  34 &   \\
+		200 & 48 &  34 &   \\
+		500 & 48 &  34 &   \\
+		1000 & 107479  &  19469 &   \\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}
+	\caption{Function evaluations}
+	\label{tab:1}
+	\center
+	\begin{tabular}{cccc}
+		\hline\noalign{\smallskip}
+		$N$ & $I_{MAM}$  & $F_{MAM}$ & $F_{SQP}$ \\
+		\hline\noalign{\smallskip}
+		100 & 48 &  34 &  \\
+		200 & 48 &  34 &  \\
+		500 & 48 &  34 &  \\
+		1000 & 107479 &  19469 &  \\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
 
@@
 Recent developments in the Multipoint Approximation Method (MAM) made it capable of solving large scale industrial optimization problems. 
-The approach uses metamodels built in the space of design variables.
-MAM solves initial problem by using approximations of the objective function and constraints.
 \Russian
+Основной особенностью MAM является тот факт, что метод solves initial problem by using approximations of the objective function and constraints. 
 В рамках данного исследования разработана параллельная версия МАМ, ориентированная на сокращение времени работы оптимизационного алгоритма (в предположении, что вычисление значений the objective function and constraints уже распараллелены).
-Проведенные эксперименты показывают приемлемое ускорение при решении задач большого размера при использовании 16 ядер.
-
+Проведенные эксперименты показывают приемлемое ускорение при решении задач большого размера при использовании 16 ядер на одном узле кластера.
 Performance is demonstrated on a benchmark example of structural optimization known as the scalable cantilevered beam.

### 2017-11-20 2017/paper_lobach.tex

%%%%%%%%%%%%%%%%%%%%%%% file template-ljm.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package ljm-auth
% for Lobachevskii Journal of Mathematics 2009/07/20
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt]{amsart}

\usepackage{ljm-auth}

\usepackage{graphicx}

\usepackage{cite}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}

\author{Strongin R.G., Gergel V.P., Barkalov K.A., Sysoyev A.V.}
\crauthor{Strongin, Gergel, Barkalov, Sysoyev} % for running heads

\tit{Generalized Parallel Computational Schemes for Time-Consuming Global Optimization}
\shorttit{Parallel Schemes for Global Optimization} %for running heads

\setcounter{page}{1}

\begin{document}

\maketit

\address{Lobachevsky State University of Nizhni Novgorod}

\email{gergel@unn.ru}

\abstract{This paper addresses computationally intensive global optimization problems, for solving of which the supercomputing systems with exaflops performance can be required. To overcome such computational complexity, the paper proposes the generalized parallel computational schemes, which may involve numerous efficient parallel algorithms of global optimization. The proposed schemes include various ways of multilevel decomposition of parallel computations to guarantee the computational efficiency of supercomputing systems with shared and distributed memory multiprocessors with thousands of processors to meet global optimization challenges.}

\notes{0}
{
%\subclass{} % 2010 Mathematical Subject Classification

\keywords{
global optimization, information-statistical theory, parallel computations, high-performance computing, supercomputing technologies
}

\thank{
This research was supported by the Russian Science Foundation, project No 16-11-10150 "Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance"
}
}

\section{Introduction}

Forthcoming exascale supercomputing systems with exaflops performance are intended to solve the computationally intensive problems with huge computational complexity. For an efficient use of the exascale supercomputers, the algorithms to solve such problems must be supreme parallel to utilize the great computational potential of millions of processors/nodes.

Undoubtedly, the global optimization problems arising in various areas of applications are computationally intensive ones, which may require the exascale supercomputing. These are the most complex problems of the theory and practice of optimal decision making. The optimized functions in these problems can be the multiextremal ones, i.e. can have numerous local optima with different values within the search domain. The multiextremality makes the optimization problems much more complicated as it requires the analysis of the entire feasible search domain. Even if the optimization problem dimensionality is small sufficiently, solving the global optimization problems requires a significant amount of computations (as few as several optimized variables can require the exascale supercomputing).

In addition, the global optimization problems are usual in the most difficult optimal decision making cases, i.e. during the computer-aided design of complex devices and systems. In the context of such problems, the performance criteria are nonlinear mostly, the search domains may be disconnected, and, above all, computing the function values may be computational intensive.

The state-of-the-art in the field of global optimization is outlined in a number of key works – see, for instance, \cite{Floudas, Locatelli, Strongin1, Pardalos, Sergeyev1, Paulavicius}. The development of the theory and methods of use of the high-performance computing for solving the computationally intensive problems of global optimization is one of the trends developing rapidly \cite{Strongin1, Ciegis, Luque, Strongin2}.

The present paper proposes generalized parallel computational schemes, which can be applied for constructing numerous efficient parallel algorithms of global optimization. The proposed schemes include various ways of multilevel decomposition of parallel computations to guarantee the computational efficiency of supercomputing systems with shared and distributed memory multiprocessors with thousands of processors to meet the global optimization challenges.

The rest of the paper has the following structure. Chapter 2 formulates the multidimensional problems of global optimization and describes an approach to the reduction of these ones to the one-dimensional optimization problems. Chapter 3 presents the generalized schemes of parallel computations for the supercomputing systems with shared and distributed memory multiprocessors. Chapter 4 presents the results of numerical experiments, which confirm the efficiency of the developed computational schemes. This is followed by the conclusions and potential ways of further research.

\section{Multidimensional problems of global optimization and their reduction}

A multidimensional global optimization problem can be stated as follows: to find a global minimum point $y^\ast \in R^N$ where the optimized function $\varphi(y)$ has the minimum value, i.e.
\begin{equation}
\varphi^\ast = \varphi(y^\ast) = min \{\varphi(y): y\in D \},
\end{equation}
where $D$ is the search domain with given boundary vectors $a$ and $b$
\[
D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.
\]
It is supposed that the objective function $\varphi(y)$ satisfies a Lipschitz condition
\begin{equation}
\vert \varphi(y') - \varphi(y'') \vert \leq L\| y' - y'' \|, \quad y', y'' \in D,
\end{equation}
where $L > 0$ is the Lipschitz constant, and $\|\ast\|$ is the Euclidian norm in the space $R^N$. Satisfying the Lipschitz condition is important as it allows estimating the global minimum on the basis of the finite number of computed values of the optimized function.

It is supposed also that the minimized function $\varphi(y)$ is defined by a computational procedure for computing the value $\varphi(y)$ at any point $y \in D$ (hereafter such computations will be called \textit{trials}). It should be pointed out that this procedure can be computationally intensive and, hence, the computational cost of solving the optimization problem (1) is determined mainly by the number of executed trials.

As it is well known (see, for example, \cite{Strongin1}), the numerical evaluation of the global optimum points of the minimized functions requires constructing a coverage of the search domain $D$. This results in a very high computational intensity of global optimization problems even in the case of small dimensionality.

The computational complexity can be reduced considerably if the computational grids resulting from covering the search domain are non-uniform, i.e. when the grid nodes are distributed denser around the global optimal point only. The construction of such non-uniform coverings may be ensured in the case of an adaptive computation, when in order to select the new trial points, all search information obtained in the course of computation (previous trial points and the values of the minimized function at these points) is used. This necessary condition complicates the computational schemes of the global optimization methods essentially as it involves the computationally intensive analysis of large amounts of the multidimensional search information. As a result, many optimization algorithms use various dimensionality reduction schemes to certain extent (see, for example, \cite{Strongin1, Strongin2, Sergeyev2}).

In the framework of the information-statistical theory, the dimensionality reduction is provided by using the Peano space-filling curves $y(x)$, which map the interval $[0, 1]$ onto a $N$-dimensional domain $D$ \cite{Strongin1, Sergeyev2, Strongin3}. As a result of such reduction, the initial multidimensional global optimization problem (1) is reduced to a one-dimensional one:
\begin{equation}
\varphi(y(x^\ast)) = \min\{\varphi(y(x)): x \in [0,1]\}.
\end{equation}
for which the function $\varphi(y(x))$ satisfies a uniform H\"older condition \cite{Strongin1}, i.e.
\begin{equation}
|\varphi(y(x')-\varphi(y(x''))| \leq H|x' - x''|^{1/N}, x', x'' \in [0,1]\},
\end{equation}
where the H\"older constant $H$ is determined as $H = 2L\sqrt{N + 3}$, $L$ is the Lipschitz constant from (2) and $N$ is the optimization problem dimension (1). A one-dimensional function $\varphi(y(x))$ is produced by reduction of the multi-dimensional function $\varphi(y)$. In order to calculate the values of $\varphi(y(x))$, the multidimensional image $y \in D$ should be computed first in accordance with the mapping $y(x)$, and then the value of the multidimensional function $\varphi(y)$ at the multidimensional point $y \in D$ should be evaluated. The resulting value $z = \varphi(y)$ is used further as the value of the one-dimensional function $\varphi(y(x))$.

\section{Multilevel parallel computational schemes for solving computationally intensive global optimization problems}

The information-statistical theory proposes a general parallelization approach for solving the global optimization problems \textit{where parallelism is provided by simultaneous computations of the values for the minimized function $\varphi(y)$ at several different points in the search domain $D$} -- see, for example, \cite{Strongin1, Strongin2, Strongin4}. Such an approach ensures the parallelism of the most computationally intensive part of the global optimization and has a general nature as it is applicable to almost any global search method for solving any kind of global optimization problems.

\subsection{General computational scheme of global optimization algorithms}

The information-statistical theory \cite{Strongin1, Strongin3} formed a basis for many efficient global optimization methods -- see, for example \cite{Strongin2, Sergeyev4, Gergel1, Gergel2, Barkalov, Gergel3, Gergel4, Gergel5, Lera}. Moreover, a general computation scheme \cite{Strongin1, Strongin2, Strongin3, Grishagin1, Grishagin2} was proposed for all developed global search algorithms. This scheme can be represented as follows.

The initial global search algorithm iteration takes place at an arbitrary point $x^1 \in (0,1)$. Let then $k, k > 1$, global search iterations have been executed. Selection of the trial point ($k + 1$) for the next iteration is subject to the following rules.

\textit{Rule}~1. Renumber the previous iteration points by subscripts in the increasing order of coordinates
\begin{equation}
0 = x_0 < x_1 < \dots < x_i < \dots < x_k < x_{k+1} = 1,
\end{equation}
the points $x_0, x_{k+1}$ were added for convenience of further description.

\textit{Rule}~2. Compute a quantity $R(i)$ for each interval $(x_{i-1}, x_i), 1 \leq i \leq k + 1$ (hereafter, these quantities will be called \textit{characteristics}).

\textit{Rule}~3. Determine the interval with the maximum characteristic
\begin{equation}
R(t) = \max_{1 \leq i \leq k + 1}R(i).
\end{equation}

\textit{Rule} 4. Execute a new trial (compute the value of the minimized function $\varphi(y(x)))$ at the point $x^{k+1}$ in the interval with the maximum characteristic from (6).

The stopping condition to terminate the optimization iterations is defined by the inequality
\begin{equation}
(x_t - x_{t-1}) \leq \varepsilon,
\end{equation}
for the interval $t, 1 \leq i \leq k + 1$ with the maximum characteristic from (6) and $\varepsilon > 0$ is the required problem solving accuracy. If the stopping condition is not satisfied, the iteration number $k$ is incremented by unity, and a new global search iteration is executed.

The computed characteristics $R(i), 1 \leq i \leq k + 1$ can be understood as the importance indicators for the intervals with respect to the global minimum point location. This provides the explanation for the interval selection scheme for executing the next trial: the point of every next global search iteration is selected in the interval with the maximum characteristic (i.e. in the interval, which is the most likely one to contain the global minimum point).

According to this general scheme, specific global search algorithms are formulated by defining the formulas to compute the characteristics $R(i), 1 \leq i \leq k+1$, and the points of the next trials in the intervals with maximum characteristics. Thus, for a multidimensional general algorithm of global search (MGAGS) \cite{Strongin1}, the interval characteristic is defined as
\begin{equation}
R(i)=m\rho_i+\frac{(z_i-z_{i-1})^2}{m\rho_i}-2 (z_i + z_{i-1}), 1 \leq i \leq k + 1,
\end{equation}
where $m$ is the H\"older constant estimate from (4), $z_i = \varphi(y(x_i)), 1 \leq i \leq k + 1$ are the computed values of the minimized function $\varphi(y(x))$ at the points of the executed iterations, and $\rho_i = \left(x_i - x_{i-1} \right)^{1/N}, 1 \leq i \leq k + 1$.

The next trial point in the case of MGAGS is computed as follows:
\begin{equation}
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{m}\right]^N,
\end{equation}
where $r, r > 1$ is a \textit{parameter} of the algorithm.

Additional information about the global optimization algorithms that can formulated in accordance with the described general computational scheme can be found in \cite{Strongin1, Strongin2}.

\subsection{Competitive parallel computational schemes for shared memory multiprocessors}

As it has been noted above, the general approach to provide the parallel computations consists in simultaneous calculating the minimized function values at several different points of the searching domain $D$. As the interval characteristics $R(i), 1 \leq i \leq k + 1$ (see Subsection 3.1) are understood as the importance indicators for the intervals with respect to the global minimum point location, additional intervals for several trials to be run in parallel should be selected subject to the following values of the characteristics.

As a result, a general scheme of the parallel algorithms of global optimization for systems with shared memory can be formulated on the basis of the scheme given in Subsection 3.1 where the following set of rules should be modified (see also \cite{Strongin1, Strongin2, Strongin4, Grishagin2}):

\textit{Rule} $3'$. Arrange the interval characteristics in the descending order
\begin{equation}
R(t_1) \geq R(t_2) \geq \dots \geq R(t_k) \geq R(t_{k+1})
\end{equation}
and select $p$ intervals with the numbers $t_j, 1 \leq j \leq p$ having the maximum values of characteristics ($p$ is the number of processors (cores) used for the parallel computing).

\textit{Rule} $4'$. Execute new trials (compute the values of the minimized function $\varphi(y(x))$) at the points $x^{k+j}, 1 \leq j \leq p$ located within the intervals with the maximum characteristics from (10).

The algorithm stopping condition (7) should be checked for all intervals where the trials are executed
\begin{equation}
\rho_{t_j} \leq \varepsilon, 1 \leq t_j \leq p.
\end{equation}
As it was mentioned above, if the stopping condition is not satisfied, the iteration number $k$ is incremented by $p$, and a new global search iteration should be executed.

In general, the described general scheme of parallel algorithms of global optimization for systems with shared memory identifies the procedure for competitive parallel computations (see Fig.~1a) where the search intervals $(x_{i-1} ,x_i), 1 \leq i \leq k + 1$ compete with each other for the access to the processors to execute next global search iterations. Taking into account the possible number of processors/cores in computational nodes with homogeneous shared memory, the described approach enables the parallelism of about $10^2$ computing units.

\begin{figure}[t]
\centering
\includegraphics[width=13.0cm]{Fig1}
\caption{Generalized computational scheme of parallel algorithms}
\label{fig:Fig1}
\end{figure}

The conditions of convergence and nonredundancy of the resulting parallel computations for the global optimization algorithms formulated in the framework of general scheme of parallel algorithms of global optimization are given in \cite{Strongin1, Strongin2, Strongin4, Grishagin2}.

\subsection{Distributed parallel computations for high-performance systems}

The next level of parallel computations in the high-performance systems consists in the use of a number of computational nodes with distributed memory. A general approach can consist in the construction of a family of linked problems based on the initial global optimization problem (1)
\begin{equation}
\overrightarrow{\Phi} = \{ \phi_l(y), y \in D, 1 \leq l \leq s \},
\end{equation}
which can be solved in parallel using different computational nodes.

Generating such a family of problems can take place during the solving of the multicriterial optimization problems -- see, for example \cite{Gergel3, Gergel4}. Another approach used widely was proposed in \cite{Strongin1, Strongin2, Strongin5, Gergel1}. This approach uses a set of mappings
\begin{equation}
Y_s(x) = \{ y^1(x), \dots , y^s(x) \},
\end{equation}
allowing constructing the necessary family of optimization problems 
\begin{equation}
\phi\bigl(y^l(x^\ast_l)\bigr) = \min \bigl\{\phi\bigl(y^l(x)\bigl): x \in [0, 1] \bigl\}, 1 \leq l \leq s,
\end{equation}
for which the global minimum points coincide with the solution of the initial problem (1), i.e. $ y^\ast = y^l(x_l^\ast), 1 \leq l \leq s$. It is important to note that a family of the one-dimensional problems $\phi(y^l(x)), 1 \leq l \leq s$ resulting from the dimensionality reduction is an information-linked one: the function values computed for some function $\phi(yl(x))$ can be used by all other problems of the family (14).

Using such a parallelization approach can result in the following general scheme of parallel algorithms of global optimization for the systems with distributed and shared memory -- see also \cite{Gergel1} and Fig.~1b.

\begin{enumerate}
\item The family of one-dimensional reduced information-linked problems (14) is distributed among the computational nodes of high-performance system. One or more problems of the family $\overrightarrow{\Phi}(y)$ from (12) may be allocated to each separate node.

\item To solve the problems, each node uses the general scheme of parallel algorithms of global optimization for the system with shared memory described in Subsection 3.2 subject to the following rules of information interchange:

\begin{itemize}
\item Before starting a new trial for any problem $\phi(y^l(x)), 1 \leq l \leq s$ at any point $x' \in [0,1]$, it is necessary to calculate the image $y' \in D$ of the point $x' \in [0,1]$ according to the mapping of $y^l(x)$ and compute the preimages of $x_i', 1 \leq i \leq s$ for each problem of the family $\overrightarrow{\Phi}(y)$. The resulting preimages $x_i', 1 \leq i \leq s$ should be sent to all involved computational nodes to eliminate repeated selection of new trial points in the intervals containing the preimages.

\item Upon the completing of any trial for any problem $\phi(y^l(x)), 1 \leq l \leq s$ at any point $x' \in [0,1]$, calculate all preimages $x_i', 1 \leq i \leq s$ again and distribute these preimages and the trial result $z' = \phi(y^l(x'))$ among all involved computational nodes to include the obtained data in the search information (5).

\item Before starting the next global search iteration, the algorithm should accept the received data and include these ones into the search information (5).
\end{itemize}
\end{enumerate}

The described general scheme of parallel global optimization algorithms enables the parallel computing for the high-performance systems, which may contain many multiprocessor (multicore) computational nodes with the distributed memory. In addition, the computers may also include Intel Xeon Phi or GPGPU-based accelerators. Taking into account the required interchange of the optimization data, the estimated number of the computational nodes with distributed memory, which the proposed scheme allows to use efficiently makes up to $10^3-10^4$ (or total up to $10^5-10^6$ computational cores).

The presented parallel computational scheme was discussed in \cite{Strongin1, Strongin2, Strongin5, Gergel1}.

\subsection{Multilevel decomposition of parallel computations for solving global optimization problems}

Further development of the parallel computations for solving the global optimization problems with the utilization of ever increasing number of processors/nodes can be ensured by integrating the computational scheme presented above with one more approach for the dimensionality reduction. This approach consists in a multistage decomposition scheme for the optimization problems \cite{Strongin2, Sergeyev4, Gergel5}. According to this scheme, a multidimensional global optimization problem can be reduced to a sequence of the nested one-dimensional problems:
\begin{equation}
\min \lbrace \varphi(y): y \in D \rbrace = \min_{a_1 \leq y_1 \leq b_1} \min_{a_2 \leq y_2 \leq b_2} ... \min_{a_N \leq y_N \leq b_N} \varphi(y).
\end{equation}

The multistage scheme (15) can be generalized for the integration with the dimensionality reduction scheme based on the Peano curves \cite{Sysoyev}. According to the generalized multistage block scheme, the vector of variables $y \in D$ of the global optimization problem (1) is considered to be a set of block variables
\[
y = (y_1, y_2, ..., y_N) = (u_1, u_2, ..., u_M)
\]
where the $i$-th block variable $u_i$ consists of the consecutive elements of the vector $y$, i.e.
\[
u_1 = (y_1, y_2, ..., y_{N_1}), \dots, 
u_M = (y_{N-N_M+1}, y_{N-N_M+2}, ..., y_N).
\]
where $N_1 + N_2 + \dots + N_M = N$.

Using new variables $u_i, 1 \leq i \leq M$, the basic relation of the multistage scheme (15) can be rewritten in the form
\begin{equation}
\min_{y \in D} \varphi(y) = \min_{u_1 \in D_1} \min_{u_2 \in D_2} ... \min_{u_M \in D_M} \varphi(y),
\end{equation}
where the subdomains $D_i, 1 \leq i \leq M$, are the projections of the initial search domain $D$ onto the subspaces corresponding to the variables $u_i,1 \leq i \leq M$. As a result, the nested subproblems of the multistage block scheme
\begin{equation}
\varphi_i(u_1,...,u_i) = \min_{u_{i+1} \in D_{i+1}} \varphi_{i+1}(u_1,...u_i,u_{i+1}), 1 \leq i \leq M - 1
\end{equation}
are the multidimensional ones and can be reduced using the Peano curves as before.

To enable the parallel computations, numerous nested optimization subproblems can be generated at once on each decomposition level for solving these ones in parallel \cite{Sergeyev7} (see Fig.~1c). The resulting number of parallel subproblems can be controlled using a predefined parallelization vector
\begin{equation}
\pi = (\pi_1, \pi_2, ..., \pi_M),
\end{equation}
where $\pi_i, 1 \leq i \leq M$, is the number of subproblems in the $(i+1)$-th decomposition level being solved in parallel. For the $M$-th level, $\pi_M$ means the number of parallel trials in the course of minimization of $\varphi_M(u_1,...,u_M) = \varphi_M(y_1,...,y_N)$ by the variable $u_M$  subject to the fixed values of $u_1, \dots, u_{M-1}$. Thus, the total number of the involved processors/cores equals to
\begin{equation}
\Pi = 1 + \sum_{i=1}^{M-1} \prod_{j=1}^i \pi_j.
\end{equation}

Using the proposed parallel implementation of the multistage block scheme finalizes constructing the generalized computational scheme of parallel global optimization algorithms. The developed approach provides the efficient use of all processors/nodes of the exascale supercomputers. Thus, for example, for 10 parallel subproblems at each stage of the multistage block scheme with 3 reduction levels (i.e. $\pi_i = 10, 1 \leq i \leq M, M = 3$) the possible number of computational units running simultaneously can be $p = \ \sim10^8-10^9$.

\section{Results of computational experiments}

The computational experiments were performed using the Lobachevsky supercomputer at the State University of Nizhny Novgorod (operating system -- CentOS 6.4, supercomputer management system - SLURM). Each computational node had two 8-core Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 GB RAM, and 3 NVIDIA Kepler K20Х graphic processors ($2\,688$ cores, 6 GB GDDR5 RAM). Intel C++ 14.0.2 was used to generate the executable program code. SDK CUDA Toolkit 6.0 was used to generate the applications for the graphic processors.

To carry out the computational experiments, the Globalizer system \cite{Sysoyev} was used. Its algorithms were based on the generalized scheme of parallel computations for solving global optimization problems described above.

The global optimization problems within the executed experiments were constructed by the GKLS-generator \cite{Gaviano}. This generator constructs the multiextremal optimization problems with predetermined properties such as the number of local minima, the global minimum point, the function value at this point, etc. In addition, this generator can construct two different types of the optimization problems, Simple (less difficult ones) and Hard (more difficult ones).

To obtain more reliable estimates of the parallelization efficiency, 100 test problems of global optimization were solved. The maximum allowed number of parallel iterations was $Kmax = 10\,000$.

The computational experiments involved the use of $p$ nodes of the Lobachevsky supercomputer, each node used three NVIDIA Kepler K20Х graphic accelerators. Thus, for $p = 32$, the number of GPU accelerators were 96, each having $2\,688$ CUDA cores (i.e. total $258\,048$ CUDA cores). Table 1 shows the Globalizer execution time, while Table 2 gives obtained parallel speedup with respect to the execution time for single computational node.

Additional computational results for the proposed generalized parallel computational scheme are given in \cite{Gergel1, Gergel2, Barkalov, Gergel3, Gergel4}. In addition, Globalizer was used also to solve various applied problems of global optimization -- see, for example, \cite{Modorskii, Gergel6}.

\begin{table}
	\caption{Average problem solving time using \textit{p} nodes}
	\label{tab:tab1}
	\center
	\begin{tabular}{|c|c|c|c|}
	\hline 
		Problem class & \parbox[c]{1.6cm}{ \centering p = 8 } & 
		\parbox[c]{1.6cm}{ \centering p = 16 } &
		\parbox[c]{1.6cm}{ \centering p = 32 } \\
	\hline 
		Simple &  2.04 & 1.50 & 0.47 \\
	\hline 
		Hard   & 11.51 & 5.53 & 0.54 \\
	\hline 
	\end{tabular}
\end{table}

\begin{table}
	\caption{Parallel speedup with respect to the execution time for single computational node}
	\label{tab:tab1}
	\center
	\begin{tabular}{|c|c|c|c|}
	\hline 
		Problem class & \parbox[c]{1.6cm}{ \centering p = 8 } & 
		\parbox[c]{1.6cm}{ \centering p = 16 } &
		\parbox[c]{1.6cm}{ \centering p = 32 } \\
	\hline 
		Simple & 25 & 34 & 109 \\
	\hline 
		Hard   & 5  & 9  & 96  \\
	\hline 
	\end{tabular}
\end{table}

\section{Conclusion}

The paper proposes a generalized computational scheme of parallel algorithms for solving global optimization problems, which require the exascale computations. The developed approach includes a competitive parallel computational scheme for the systems with shared memory, the distributed parallel computations based on using the multiple Peano curves for the dimensionality reduction, and the multilevel decomposition of the parallel computations for the exascale supercomputers. In addition, the supercomputers may include also Intel Xeon Phi or GPGPU-based accelerators. The proposed schemes ensure the computational efficiency of the supercomputing systems with shared and distributed memory with thousands of processors to address the global optimization challenges.

A number of development trends are relevant to evolve the research. Additional theoretical analysis of the proposed parallel computation scheme should be evaluated. Also, a spectrum of parallel global optimization algorithms, which can be implemented in the framework of the proposed computational scheme, should be studied. Besides, additional computational experiments involving a larger number of computational units should be carried out to evaluate the efficiency of parallel computations for solving computationally intensive global optimization problems adequately.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{thebibliography}{20}

\bibitem{Floudas}
C.~A.~Floudas and M.~P.~Pardalos, \textit{Recent advances in global optimization} (Princeton University Press, 2016).

\bibitem{Locatelli}
M.~Locatelli and F.~Schoen, \textit{Global optimization: theory, algorithms and applications} (SIAM, 2013).

\bibitem{Strongin1}
R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global optimization with non-convex constraints. Sequential and parallel algorithms} (Kluwer Academic Publishers, Dordrecht, 2000, 2nd ed. 2013, 3rd ed. 2014).

\bibitem{Pardalos}
P.~M.~Pardalos, A.~A.~Zhigljavsky and J.~\v{Z}ilinskas \textit{Advances in stochastic and deterministic global optimization} (Springer, 2016).

\bibitem{Sergeyev1}
Y.~D.~Sergeyev and D.~E.~Kvasov, \textit{Deterministic Global Optimization. An Introduction to the Diagonal Approach}  (Springer Briefs in Optimization, Springer, 2017).

\bibitem{Paulavicius}
R.~Paulavi\v{c}ius, J.~\v{Z}ilinskas, \textit{Simplicial Global Optimization} (Springer Briefs in Optimization. Springer, 2014).

\bibitem{Ciegis}
R.~\v{C}iegis, D.~Henty, B.~K\r{a}gstr\"om and J.~\v{Z}ilinskas, \textit{Parallel scientific computing and optimization: advances and applications}  (Springer, 2009).

\bibitem{Luque}
G.~Luque and E.~Alba, \textit{Parallel genetic algorithms. Theory and real world applications} (Springer-Verlag, Berlin, 2011).

\bibitem{Strongin2}
R.~G.~Strongin, V.~P.~Gergel, V.~A.~Grishagin and K.~A.~Barkalov, \textit{Parallel computations for global optimization problems} (Moscow State University, Moscow, 2013) [In Russian].

\bibitem{Sergeyev2}
Ya.~D.~Sergeyev, R.~G.~Strongin, D.~Lera, \textit{Introduction to Global Optimization Exploiting Space-Filling Curves} (Springer Briefs in Optimization, Springer, 2013).

\bibitem{Strongin3}
R.~G.~Strongin, \textit{Numerical Methods in Multiextremal Problems (Information-Statistical Algorithms)} (Nauka, Moscow, 1978) [In Russian].

\bibitem{Strongin4}
R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global multidimensional optimization on parallel computer}, Parallel Computing, \textbf{18} (11), 1259--1273, (1992).

\bibitem{Sergeyev4}
Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel asynchronous global search and the nested optimization scheme}, J. Comput. Anal. Appl., \textbf{3} (2), 123--145, (2001).

\bibitem{Gergel1}
V.~P.~Gergel and S.~V.~Sidorov, \textit{A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems}, V. Malyshkin (Ed.): PaCT 2015, LNCS 9251, 505--515. Springer-Verlag Berlin Heidelberg ,(2015).

\bibitem{Gergel2}
V.~Gergel, \textit{An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems}, 2nd International Conference on Mathematics and Computers in Sciences and in Industry, 13--18, (2015).

\bibitem{Barkalov}
K.~Barkalov, V.~Gergel and I.~Lebedev, \textit{Solving global optimization problems on GPU cluster}, In: Simos, T.E. (ed.) ICNAAM 2015. AIP Conference Proceedings, 1738, art. no. 400006, (2016).

\bibitem{Gergel3}
V.~Gergel and E.~Kozinov, \textit{Efficient methods of multicriterial optimization based on the intensive use of search information}, Springer Proceedings in Mathematics and Statistics, 197, 27--45, (2017).

\bibitem{Gergel4}
V.~Gergel and E.~Kozinov, \textit{Parallel computing for time-consuming multicriterial optimization problems}, Lecture Notes in Computer Science, 10421, 446--458, (2017).

\bibitem{Gergel5}
V.~Gergel, V.~Grishagin and A.~Gergel, \textit{Adaptive nested optimization scheme for multidimensional global search}, Journal of Global Optimization, \textbf{66} (1), 35--51, (2016).

\bibitem{Lera}
D.~Lera, Y.~D.~Sergeyev, \textit{Lipschitz and Holder global optimization using space-filling curves}, Appl. Numer. Math, \textbf{60} (1-2), 115--129, (2010).

\bibitem{Grishagin1}
V~A.~Grishagin, \textit{On convergence conditions for a class of global search algorithms}, Proceedings of the 3-rd All-Union seminar "Numerical methods of nonlinear programming", Kharkov, 82--84, (1979) [In Russian].

\bibitem{Grishagin2}
V.~A.~Grishagin, Y.~D.~Sergeyev and R.~G.~Strongin, \textit{Parallel characteristic algorithms for solving problems of global optimization}, Journal of Global Optimization, \textbf{10} (2), 185--206, (1997).

\bibitem{Strongin5}
R.~G.~Strongin, \textit{Algorithms for Multi-extremal Mathematical Programming Problems Employing the Set of Joint Space-filling Curves}, Journal of Global Optimization, \textbf{2} (4), 357--378, (1992).

\bibitem{Sysoyev}
A.~Sysoyev, K.~Barkalov, V.~Sovrasov, I.~Lebedev and V.~Gergel, \textit{Globalizer –- A parallel software system for solving global optimization problems}, Lecture Notes in Computer Science 10421, 492--499, (2017).

\bibitem{Sergeyev7}
Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel Asynchronous Global Search and the Nested Optimization Scheme}, J. Comput. Anal. Appl., \textbf{3} (2), 123--145, (2001).

\bibitem{Gaviano}
M.~Gaviano, D.~Lera, D.~E.~Kvasov and Ya.~D.~Sergeyev, \textit{Software for generation of classes of test functions with known local and global minima for global optimization}, ACM Trans. Math. Software, \textbf{29}, 469--480, (2003).

\bibitem{Modorskii}
V.~Y.~Modorskii, D.~F.~Gaynutdinova, V.~P.~Gergel and K.~A.~Barkalov, \textit{Optimization in design of scientific products for purposes of cavitation problems}, AIP Conference Proceedings, 1738, art. no. 400013, (2016).

\bibitem{Gergel6}
V.~P.~Gergel, M.~I.~Kuzmin, N.~A.~Solovyov and V.~A.~Grishagin, \textit{Recognition of surface defects of cold-rolling sheets based on method of localities}, International Review of Automatic Control, \textbf{8} (1), 51--55, (2015).

\end{thebibliography}

\end{document}

### 2017-11-21 2017/KAZAN_LJM/paper_lobach.tex
min -> \min + красивые кавычки

--- 2017-11-20 2017/paper_lobach.tex
@@
 \thank{
-This research was supported by the Russian Science Foundation, project No 16-11-10150 "Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance"
+This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''
 }
@@
 \begin{equation}
-\varphi^\ast = \varphi(y^\ast) = min \{\varphi(y): y\in D \},
+\varphi^\ast = \varphi(y^\ast) = \min \{\varphi(y): y\in D \},
 \end{equation}
@@
 	\caption{Parallel speedup with respect to the execution time for single computational node}
-	\label{tab:tab1}
+	\label{tab:tab2}
 	\center
@@
 \bibitem{Grishagin1}
-V~A.~Grishagin, \textit{On convergence conditions for a class of global search algorithms}, Proceedings of the 3-rd All-Union seminar "Numerical methods of nonlinear programming", Kharkov, 82--84, (1979) [In Russian].
+V~A.~Grishagin, \textit{On convergence conditions for a class of global search algorithms}, Proceedings of the 3-rd All-Union seminar ``Numerical methods of nonlinear programming'', Kharkov, 82--84, (1979) [In Russian].
 

### 2017-11-21 2018/PAVT_Toropov/PAVT2018_MAM.tex
Iteration number corrected
Translate to English

--- 2017-11-13 2018/PAVT_Toropov/PAVT2018_MAM.tex
@@
 \usepackage{url}
+
+\usepackage[T2A]{fontenc}
+\usepackage[utf8x]{inputenc}
+\usepackage[russian,english]{babel}
 
@@
 \begin{abstract}
-TBD
+The paper presents a new development in the Multipoint Approximation Method (MAM) that makes it capable of handling large scale problems.
+The approach relies on approximations built in the space of design variables within the iterative trust-region-based framework of MAM.
+With the purpose of solving the problems of high dimensionality in a reasonable time, a parallel variant of the Multipoint Approximation Method (PMAM) has been developed.
+It was supposed that the values of the objective function and of the constraints were computed using the distributed memory (on several cluster nodes), whereas the optimization module run on a single node using the shared memory.
+The numerical experiments have been carried out on a benchmark example of structural optimization.
 
@@
 \section{Introduction}
-TBD
 
-In problems with a large (in the order of hundreds) number of design variables, the multipoint approximation method (MAM \cite{Toropov1989,Toropov1992,ToropovFilatov1993}) proved to be efficient, e.g.
-in turbomachinery applications \cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010} .
+\Russian
+In the present paper, the multipoint approximation method (MAM \cite{Toropov1989,Toropov1992,ToropovFilatov1993}) and its application to the large scale optimization problems are considered.
+In problems with a large (in the order of hundreds) number of design variables MAM proved to be efficient, e.g.
+in turbomachinery applications \cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010}.
 This method is an iterative optimization technique based on mid-range approximations built in trust regions.
@@
 
-TBD
+This paper presents a Parallel Multipoint Approximation Method that makes it capable of handling problems with the number of design variables in the order of thousands.
+The parallel variant of the algorithm (with the use of the shared memory) has been developed with the purpose of minimizing the worktime of the part of the algorithm related to constructing the approximation and solving the approximated problem and not related to computing the values of the objective function and  constraints.
+The processes of computing the values of the objective function and constraints (with the use of the
+distributed memory) are supposed to be parallelized already.
 
@@
 The main loop of the MAM is organized as follows.
+\\
 
-Algorithm (MAM).
+\textbf{Algorithm (MAM).}
 \begin{enumerate}
@@
 
-\section{Parallel Algorithm}
-\label{sec:par_alg}
+\section{Parallel Multipoint Approximation Method}\label{sec:par_alg}
+
+Let us consider possible methods of parallelization, which could be applied for the considered problems.
+
+First, one can parallelize the computing of the functions describing the optimized object.
+This way is obvious as well as necessary one since in the industrial design optimization problems the computing of even a single function value might take several hours.
+However, this method is a specific one for each particular problem.
+Here the issues of computations are addressed at the level of the application software, in which the industrial modeling is performed (e.g.
+Ansys, OpenFOAM, etc.).
+
+Second, one can correct the algorithm with the purpose of parallel computing several values of the objective function and constraints at different points of the search domain.
+According to the MAM rules, in design of experiments $P$
+sampling points are formed in current trust region at every iteration.
+The function values at these points can be computed on different processors (nodes) in parallel.
+The above corresponds to the parallelization of Step 2a of the algorithm with the use of the distributed memory.
+The number of sampling points generated within each iteration may be set equal to $P = k\cdot NP$ number, where $NP$ is a number of available processors (or nodes) and $k \geq 1$.
+In  terms of the time, the latter will be equivalent to $NP$ function evaluations per step.
+This method has been implemented successfully in \cite{PolynkinToropov2008} and has demonstrated a good efficiency since at the number of the design variables of the order of $10^2$ major time is consumed by the function evaluations, the work of the MAM itself (in the sequential regime) introduces minor overhead.
+
+However, when the number of variables becomes of the order of $10^3$, the work of the sequential part of MAM begins to affect the total problem solving time essentially (assuming the time of computing the objective function
+values remains constant).
+Thus, for the problem considered in Section 4 the time of execution of a single iteration of the method increased more than 2000 times (from 4 up to 9000 seconds) when increasing the number of variables from 100 up to  1000.
+Therefore, another approach to the parallelization of the algorithm has been applied within the framework of the
+present study: namely, the computational rules of MAM  providing the construction of the approximating, the solving of the approximated problem, and the choice of the next trust region (Steps 2b, 2c, and 2d of the algorithm) have been parallelized.
+
+In order to find the most time-consuming parts of the sequential program developed earlier, we have applied Intel VTune Amplifier XE.
+The performed analysis has shown the matrix multiplication and solving SLAE when constructing the approximations as well as solving the approximating problem by SQP method to be the most time-consuming operations in the execution of MAM.
+The matrix multiplication and solving SLAE are standard operations implemented in many high-performance libraries.
+Here we used the corresponding parallel methods from Intel MKL library.
+We have parallelized the SQP method ourselves using OpenMP.
 
@@
 
+Next, let us compare the worktime of the sequential and parallel algorithms when solving the large scale problems. 
+The dimensionality of the problem being solved $N$ was varied from 100 up to 1000 design variables that corresponds to 50 to 500 segments of the cantilevered beam.
+The number of points in a trust region used to build the approximations was $1.5 N$.
+Both algorithms was run on a single node of the cluster (the specifications of the node are listed below), the parallel algorithm employed all 16 processors cores available.
+Since the time of computing of the objective
+function and constraints in the test problem being solved was negligible, these ones were computed on the same node.
+
+Table 1 reflects the worktime (in seconds) of the sequential algorithm and of the parallel one subject to the number of variables.
+In Table 2, the number of MAM iterations $I_{MAM}$ as well as the number of the function evaluations $F_{MAM}$ required for solving the problem are presented.
+For comparison, the number of the function value computations, which would be required to solve the initial (non-approximated) problem by SQP method $F_{SQP}$ is given also.
+In all conducted experiments, the objective function values in the sequential and parallel versions of the algorithm were the same (up to the accuracy of the computational error) and differed from the solution obtained by SQP negligibly.
+The reduction of the number of the function evaluations required to solve the problem using MAM as compared to the use of SQP was demonstrated
+visibly.
+
+\begin{table}
+	\caption{Time and speedup}
+	\label{tab:1}
+	\center
+	\begin{tabular}{cccc}
+		\hline\noalign{\smallskip}
+		$N$ & MAM time & PMAM time & Speedup \\
+		\hline\noalign{\smallskip}
+		100 & 48 &  34 &   \\
+		200 & 48 &  34 &   \\
+		500 & 48 &  34 &   \\
+		1000 & 107479  &  19469 &   \\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}
+	\caption{Function evaluations}
+	\label{tab:2}
+	\center
+	\begin{tabular}{cccc}
+		\hline\noalign{\smallskip}
+		$N$ & $I_{MAM}$  & $F_{MAM}$ & $F_{SQP}$ \\
+		\hline\noalign{\smallskip}
+		100 & 10 &  2201 & 9901 \\
+		200 & 10 &  4402 & 28143 \\
+		500 &  &   &  \\
+		1000 &  &   &  \\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+Computational experiments were carried out on a high-performance cluster of Lobachevsky State University of Nizhny Novgorod.
+The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.
+The CPU had 8 cores, i.e.
+a total of 16 physical cores were available in the node.
+MS Visual Studio 15 and Intel Fortran Compiler were used to implement the algorithm.
+
 \section{Conclusions}
+
+Recent developments in the Multipoint Approximation Method (MAM) made it capable of solving large scale industrial optimization problems.
+The fact that MAM solves the initial problem by using the approximations of the objective function and constraints is the primary distinctive feature of the method.
+Within the framework of the present study, the parallel version of MAM oriented onto the reduction of the worktime of the optimization algorithm (assuming that computing the values of the objective function and constraints has been parallelized already) has been developed.
+The performed experiments have demonstrated an acceptable speedup when solving the large scale problems employing 16 cores on a single cluster node.
+The performance was demonstrated on a benchmark example of structural
+optimization known as the scalable cantilevered beam.
 

### 2017-11-22 2018/PAVT_Toropov/PAVT2018_MAM.tex
Table 1

@@
 
-\usepackage[T2A]{fontenc}
-\usepackage[utf8x]{inputenc}
-\usepackage[russian,english]{babel}
+%\usepackage[T2A]{fontenc}
+%\usepackage[utf8x]{inputenc}
+%\usepackage[russian,english]{babel}
 
@@
 
-\Russian
 In the present paper, the multipoint approximation method (MAM \cite{Toropov1989,Toropov1992,ToropovFilatov1993}) and its application to the large scale optimization problems are considered.
@@
 values remains constant).
-Thus, for the problem considered in Section 4 the time of execution of a single iteration of the method increased more than 2000 times (from 4 up to 9000 seconds) when increasing the number of variables from 100 up to  1000.
+Thus, for the problem considered in Section 4 the time of execution of a single iteration of the method increased more than 4000 times (from 1.5 up to 6000 seconds) when increasing the number of variables from 100 up to  1000.
 Therefore, another approach to the parallelization of the algorithm has been applied within the framework of the
@@
 		\hline\noalign{\smallskip}
-		$N$ & MAM time & PMAM time & Speedup \\
+		$N$ & $T_{MAM}$ & $T_{PMAM}$ & Speedup \\
 		\hline\noalign{\smallskip}
-		100 & 48 &  34 &   \\
-		200 & 48 &  34 &   \\
-		500 & 48 &  34 &   \\
-		1000 & 107479  &  19469 &   \\
+		100 & 15.4 & 11.9 & 1.3  \\
+		200 & 167 &  89 &  1.9 \\
+		500 & 5151 &  1520 & 3.4  \\
+		1000 & 67674 &  13771 & 4.9  \\
 		\noalign{\smallskip}\hline
@@
 		200 & 10 &  4402 & 28143 \\
-		500 &  &   &  \\
-		1000 &  &   &  \\
+		500 &  &   &  115734 \\
+		1000 &  &   & 305308 \\
 		\noalign{\smallskip}\hline

### 2017-11-23 2018/PAVT_Toropov/PAVT2018_MAM.tex
References updated

@@
 \begin{matrix}%[pos]{spalten}
-			b=& 3.246& 3.224& 3.202& 3.179& 3.156& 3.133& 3.109& 3.085& 3.061& 3.036 \\  
-			  & 3.011& 2.985& 2.959& 2.933& 2.905& 2.878& 2.850& 2.821& 2.792& 2.762 \\
-			  & 2.731& 2.700& 2.668& 2.635& 2.602& 2.567& 2.532& 2.495& 2.458& 2.419 \\  
-			  & 2.379& 2.338& 2.295& 2.250& 2.204& 2.156& 2.105& 2.052& 1.996& 1.936 \\  
-			  & 1.873& 1.805& 1.732& 1.651& 1.562& 1.462& 1.345& 1.196& 1.023& 1.000 \\ 	
+			b=& \left[3.246, \right.& 3.224,& 3.202,& 3.179,& 3.156,& 3.133,& 3.109,& 3.085,& 3.061,& 3.036,  \\  
+			  & 3.011,& 2.985,& 2.959,& 2.933,& 2.905,& 2.878,& 2.850,& 2.821,& 2.792,& 2.762, \\
+			  & 2.731,& 2.700,& 2.668,& 2.635,& 2.602,& 2.567,& 2.532,& 2.495,& 2.458,& 2.419, \\  
+			  & 2.379,& 2.338,& 2.295,& 2.250,& 2.204,& 2.156,& 2.105,& 2.052,& 1.996,& 1.936, \\  
+			  & 1.873,& 1.805,& 1.732,& 1.651,& 1.562,& 1.462,& 1.345,& 1.196,& 1.023,& \left. 1.000\right] 	
 \end{matrix}
@@
 \begin{matrix}
-	h=& 64.919&64.480&64.033&63.581&63.122&62.656&62.184&61.703&61.216&60.720  \\
-		& 60.216&59.703&59.182&58.651&58.110&57.558&56.996&56.423&55.837&55.239  \\
-		& 54.628&54.003&53.363&52.707&52.035&51.344&50.635&49.906&49.155&48.381  \\
-		& 47.582&46.754&45.897&45.008&44.082&43.116&42.105&41.041&39.919&38.729  \\
-		& 37.462&36.102&34.630&33.020&31.240&29.241&26.910&23.919&20.465&14.639  \\
+	h=& \left[64.919,\right.&64.480,&64.033,&63.581,&63.122,&62.656,&62.184,&61.703,&61.216,&60.720  \\
+		& 60.216,&59.703,&59.182,&58.651,&58.110,&57.558,&56.996,&56.423,&55.837,&55.239  \\
+		& 54.628,&54.003,&53.363,&52.707,&52.035,&51.344,&50.635,&49.906,&49.155,&48.381  \\
+		& 47.582,&46.754,&45.897,&45.008,&44.082,&43.116,&42.105,&41.041,&39.919,&38.729  \\
+		& 37.462,&36.102,&34.630,&33.020,&31.240,&29.241,&26.910,&23.919,&20.465,&\left.14.639 \right]
 \end{matrix}
@@
 \begin{matrix}
-	b=&	3.238&3.204&3.202&3.170&3.176&3.128&3.108&3.089&3.057&3.003 \\  
-		& 3.027&2.986&2.959&2.952&2.885&2.855&2.864&2.847&2.803&2.762 \\  
-		& 2.737&2.722&2.630&2.645&2.590&2.558&2.547&2.480&2.693&2.391 \\ 
-		& 2.368&2.310&2.307&2.227&2.176&2.149&2.106&2.016&2.007&1.925 \\  
-		& 1.864&1.843&1.758&1.635&1.582&1.934&1.332&1.173&1.026&2.419 \\ 
+	b=&	\left[3.238,\right.&3.204,&3.202,&3.170,&3.176,&3.128,&3.108,&3.089,&3.057,&3.003 \\  
+		& 3.027,&2.986,&2.959,&2.952,&2.885,&2.855,&2.864,&2.847,&2.803,&2.762 \\  
+		& 2.737,&2.722,&2.630,&2.645,&2.590,&2.558,&2.547,&2.480,&2.693,&2.391 \\ 
+		& 2.368,&2.310,&2.307,&2.227,&2.176,&2.149,&2.106,&2.016,&2.007,&1.925 \\  
+		& 1.864,&1.843,&1.758,&1.635,&1.582,&1.934,&1.332,&1.173,&1.026,&\left.2.419\right]
 \end{matrix}
@@
 \begin{matrix}
-	h=& 64.764&64.083&64.036&63.392&63.518&62.566&62.153&61.772&61.149& 60.054  \\
-		& 60.534&59.728&59.182&59.033&57.695&57.105&57.274&56.943&56.057&55.233  \\
-		& 54.747&54.433&52.590&52.902&51.802&51.150&50.943&49.597&51.567&47.811  \\
-		& 47.373&46.190&46.140&44.543&43.525&42.991&42.119&40.308&40.141&38.506  \\
-		& 37.268&36.839&35.153&32.702&31.615&28.304&26.642&23.447&20.536& 9.417
-
+	h=& \left[64.764,\right.&64.083,&64.036,&63.392,&63.518,&62.566,&62.153,&61.772,&61.149,& 60.054  \\
+		& 60.534,&59.728,&59.182,&59.033,&57.695,&57.105,&57.274,&56.943,&56.057,&55.233  \\
+		& 54.747,&54.433,&52.590,&52.902,&51.802,&51.150,&50.943,&49.597,&51.567,&47.811  \\
+		& 47.373,&46.190,&46.140,&44.543,&43.525,&42.991,&42.119,&40.308,&40.141,&38.506  \\
+		& 37.268,&36.839,&35.153,&32.702,&31.615,&28.304,&26.642,&23.447,&20.536,& \left.9.417\right]
 \end{matrix}
@@
 
-\bibliography{bibliography}{}
-\bibliographystyle{nature}
+\begin{thebibliography}{10}
+
+\bibitem{Toropov1989}
+Toropov, V.:
+\newblock Simulation approach to structural optimization.
+\newblock Structural Optimization \textbf{1}, 37--46 (1989)
+
+\bibitem{Toropov1992}
+Toropov, V.:
+\newblock Multipoint approximation method in optimization problems with expensive function values.
+\newblock In: Sydow, A. (ed.) Proceedings of the 4th international symposium on Systems analysis and simulation, 
+pp. 207--212, Elsevier,  (1992)
+
+\bibitem{ToropovFilatov1993}
+Toropov, V., Filatov, A., Polynkin, A.:
+\newblock Multiparameter structural optimization using fem and multipoint explicit approximations.
+\newblock Structural Optimization \textbf{6}, 7--14 (1993)
+
+\bibitem{ShahparPolynkinToropov2008}
+Shahpar, S., Polynkin, A., Toropov, V.:
+\newblock Large scale optimization of transonic axial compressor rotor blades.
+\newblock In:  49th AIAA/ASME/ASCE/AHS/ASC Structures, Structural Dynamics, and Materials Conference, AIAA, art. no. 2008-2056
+ (2008)
+
+\bibitem{PolynkinToropovShahpar2008}
+Polynkin, A., Toropov, V., Shahpar, S.:
+\newblock Design optimization of aircraft engine components.
+\newblock In: Proc. of 7th ASMO UK / ISSMO Conf. on Engineering Design Optimization. Process and Product Improvement (2008)
+
+\bibitem{PolynkinToropovShahpar2010}
+Polynkin, A., Toropov, V., Shahpar, S.:
+\newblock Multidisciplinary optimization of turbomachinery based on metamodel built by genetic programming.
+\newblock In: 13th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference (2010)
+
+\bibitem{KeulenToropovMarkine1996}
+van Keulen, F., Toropov, V., Markine, V.:
+\newblock Recent refinements in the multi-point approximation method in conjunction with adaptive mesh refinement.
+\newblock In: McCarthy, J.M. (ed.) Proc. of ASME Design Engineering Technical Conferences and Computers
+  in Engineering Conf., 1--12, ASME, Irvine CA (1996)
+
+\bibitem{ToropovKeulenMarkine1996}
+Toropov, V., van Keulen, F., Markine, V., de~Boer, H.:
+\newblock Refinements in the multi-point approximation method to reduce the effects of noisy responses.
+\newblock In: 6th AIAA/NASA/ISSMO Symp. on Multidisciplinary Analysis and Optimization,  pp. 941--951, AIAA (1996)
+
+\bibitem{ToropovMarkineHolden1999}
+Toropov, V., Markine, V., Holden, C.:
+\newblock Use of mid-range approximations for optimization problems with functions of domain-dependent calculability.
+\newblock In: 3rd ISSMO/UBCAD/UB/AIAA World Congress of Structural and Multidisciplinary Optimization (1999).
+
+\bibitem{ToropovMarkine1996}
+Toropov, V., Markine, V.:
+\newblock The use of simplified numerical models as mid-range approximations.
+\newblock In: 6th AIAA/NASA/ISSMO Symp. on Multidisciplinary Analysis and Optimization, pp. 952--958 (1996).
+
+\bibitem{ToropovAlvarez1998}
+Toropov, V., Alvarez, L.:
+\newblock Creation of multipoint approximations using genetic programming.
+\newblock In: Parmee, I.C. (ed.) Adaptive Computing in Design and Manufacture, 3rd Int. Conf.,
+   pp. 21--24, PEDC, Dartington (1998)
+
+\bibitem{PolynkinToropov2012}
+Polynkin, A. Toropov, V.:
+\newblock Mid-range metamodel assembly building based on linear regression for large scale optimization problems.
+\newblock Structural and Multidisciplinary Optimization \textbf{45}(4), 515--527 (2012)
+
+\bibitem{LancasterSalkauskas1981}
+Lancaster, P. Salkauskas, K.:
+\newblock Surfaces generated by moving least squares methods.
+\newblock Mathematics of Computation \textbf{87}, 141--158 (1981)
+
+\bibitem{Liszka1984}
+Liszka, T.:
+\newblock An interpolation method for an irregular net of nodes.
+\newblock Int. J. Num. Meth. \textbf{20}, 1599--1612 (1984)
+
+\bibitem{ChoiYounYang2001}
+Choi, K., Youn, B., Yang, R.-J.:
+\newblock Moving least squares method for reliability-based design optimization.
+\newblock In: 4th World Congress of Structural and Multidisciplinary Optimization, Dalian, China (2001)
+
+\bibitem{ToropovSchrammSahaiJones2005}
+Toropov, V., Schramm, U., Sahai, A., Jones, R., Zeguer, T.:
+\newblock Design optimization and stochastic analysis based on the moving least squares method.
+\newblock In: Herskovits, J., Mazorche, S., Canelas, A. (eds.) 6th World Congress of Structural and Multidisciplinary Optimization, art. no. 9412 (2005)
+
+\bibitem{PolynkinToropov2010}
+Polynkin, A. Toropov, V.:
+\newblock Recognition of design variable inter-dependencies using  cross-validated moving least-squares method.
+\newblock In: Proceedings of the 51st AIAA/ASME/ASCE/AHS/ASC Structures, Structural Dynamics, and Materials Conference, AIAA, art. no. 2010-2985 (2010)
+
+\bibitem{OllarToropovJones2014}
+Ollar, J., Toropov, V., Jones. R.:
+\newblock  Mid-range approximations in sub-spaces for MDO problems with disparate discipline attributes.
+\newblock In: 15th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference, AIAA, art. no. 2014-2437 (2014)
+
+\bibitem{KeulenToropov1997}
+van Keulen, F. Toropov, V.:
+\newblock New developments in structural optimization using adaptive mesh refinement and multi-point approximations.
+\newblock Engineering Optimization \textbf{29}, 217--234 (1997)
+
+\bibitem{PolynkinToropov2008}
+Polynkin, A., Toropov, V., Shahpar, S.:
+\newblock Adaptive and parallel capabilities in the multipoint approximation  method.
+\newblock In: 12th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference, AIAA, art. no. 2008-5803 (2008)
+
+\bibitem{Vanderpllaats2001}
+Vanderplaats, G.:
+\newblock Multidiscipline Design Optimization. 
+\newblock Vanderplaats Research \& Development, Inc., Colorado Springs, (2001)
+
+\end{thebibliography}
 

### 2017-11-24 2018/PAVT_Toropov/PAVT2018_MAM.tex
Pred-final

@@
       \min\limits_{a_i^k \le x_i ^k\le b_i}\widetilde{F}^k_0(x) \\
-      s.t.\; \widetilde{F}^k_j( x) \le 1,\; j=1,\dots ,M,
+      s.t.\; \widetilde{F}^k_j( x) \le 1,\; j=1,\dots ,M.
       \end{array}
@@
 \end{displaymath}
-The moment of inertia of a segment is $I_i=\frac{b_i h_i^3}{12}$ and the bending moment at the left end is $M_i=P[L+l_i- \sum_{j=1}^{i}  l_j ]$.
+The moment of inertia of a segment is $I_i={b_i h_i^3}/{12}$ and the bending moment at the left end is $M_i=P[L+l_i- \sum_{j=1}^{i}  l_j ]$.
 The maximum bending stress in the segment $i$ is then given by the following formula:
@@
     5.0 \le h_i \le 100.0, \\
-    y_N\le 2/5, \\
+    y_N\le 2.5, \\
     \sigma_i \le \bar{\sigma}=14000, \\
@@
 
-MAM obtained the solution $V = 63935.360\; cm^3$, using 1500 function evaluations (as compared to almost 8000 evaluations used by SQP).
-The number of points in a trust region used to build approximations was 150.
+MAM obtained the solution $V = 63935.360\; cm^3$, using 2201 function evaluations (as compared to almost 10000 evaluations used by SQP).
+The number of points in a trust region used to build approximations was 200.
 The optimal values of design variables obtained by MAM are given below:
@@
 The dimensionality of the problem being solved $N$ was varied from 100 up to 1000 design variables that corresponds to 50 to 500 segments of the cantilevered beam.
-The number of points in a trust region used to build the approximations was $1.5 N$.
+The number of points in a trust region used to build the approximations was $2 N$.
 Both algorithms was run on a single node of the cluster (the specifications of the node are listed below), the parallel algorithm employed all 16 processors cores available.
@@
 function and constraints in the test problem being solved was negligible, these ones were computed on the same node.
-
-Table 1 reflects the worktime (in seconds) of the sequential algorithm and of the parallel one subject to the number of variables.
-In Table 2, the number of MAM iterations $I_{MAM}$ as well as the number of the function evaluations $F_{MAM}$ required for solving the problem are presented.
-For comparison, the number of the function value computations, which would be required to solve the initial (non-approximated) problem by SQP method $F_{SQP}$ is given also.
-In all conducted experiments, the objective function values in the sequential and parallel versions of the algorithm were the same (up to the accuracy of the computational error) and differed from the solution obtained by SQP negligibly.
-The reduction of the number of the function evaluations required to solve the problem using MAM as compared to the use of SQP was demonstrated
-visibly.
 
@@
 		200 & 167 &  89 &  1.9 \\
-		500 & 5151 &  1520 & 3.4  \\
+		400 & 2476 &  837 &  3.0 \\
+%		500 & 5151 &  1520 & 3.4  \\
+		600 & 9678 &  2777 &  3.5 \\
+		800 & 27826 &  6948 &  4.0 \\
 		1000 & 67674 &  13771 & 4.9  \\
@@
 		200 & 10 &  4402 & 28143 \\
-		500 &  &   &  115734 \\
-		1000 &  &   & 305308 \\
+		400 & 11 &  9599 &  84211 \\
+%		500 &  &   &  115734 \\
+		600 & 10 &  13200 &  146046 \\
+		800 & 10 &  17600 &  221079 \\
+		1000 & 10 &  22000 & 305308 \\
 		\noalign{\smallskip}\hline
@@
 \end{table}
+
+Table 1 reflects the worktime (in seconds) of the sequential algorithm and of the parallel one subject to the number of variables.
+In Table 2, the number of MAM iterations $I_{MAM}$ as well as the number of the function evaluations $F_{MAM}$ required for solving the problem are presented.
+For comparison, the number of the function value computations, which would be required to solve the initial (non-approximated) problem by SQP method $F_{SQP}$ is given also.
+In all conducted experiments, the objective function values in the sequential and parallel versions of the algorithm were the same (up to the accuracy of the computational error) and differed from the solution obtained by SQP negligibly.
+The reduction of the number of the function evaluations required to solve the problem using MAM as compared to the use of SQP was demonstrated
+visibly.
 

### 2017-12-01 2018/PAVT_Toropov/PAVT2018_MAM.tex
minor corrections

@@
 Evgeny A. Kozinov%(ORCID 0000-0001-6776-0096)
-$^1$ \and
+$^1$ \and \\
 Vassili V. Toropov%(ORCID 0000-0002-7017-5055)
@@
 In  terms of the time, the latter will be equivalent to $NP$ function evaluations per step.
-This method has been implemented successfully in \cite{PolynkinToropov2008} and has demonstrated a good efficiency since at the number of the design variables of the order of $10^2$ major time is consumed by the function evaluations, the work of the MAM itself (in the sequential regime) introduces minor overhead.
+This method has been implemented successfully in \cite{PolynkinToropov2008} and has demonstrated a good efficiency since at the number of the design variables of the order of 100 major time is consumed by the function evaluations, the work of the MAM itself (in the sequential regime) introduces minor overhead.
 
-However, when the number of variables becomes of the order of $10^3$, the work of the sequential part of MAM begins to affect the total problem solving time essentially (assuming the time of computing the objective function
+However, when the number of variables becomes of the order of 1000, the work of the sequential part of MAM begins to affect the total problem solving time essentially (assuming the time of computing the objective function
 values remains constant).

### 2017-12-12 2018/LION/LION2018.tex
references are updated

@@
 
-\bibliography{bibliography}{}
-\bibliographystyle{nature}
+%\bibliography{bibliography}{}
+%\bibliographystyle{nature}
+
+\begin{thebibliography}{10}
+
+\bibitem{RockafellarUryasev2000}
+Rockafellar, R., Uryasev, S.:
+\newblock Optimization of conditional value-at-risk.
+\newblock The Journal of Risk \textbf{2}(3), 21--41 (2000)
+
+\bibitem{SchillingsSchulz2015}
+Schillings, C., Schulz, V.:
+\newblock On the influence of robustness measures on shape optimization with stochastic uncertainties.
+\newblock Optimization and Engineering \textbf{16}(2), 347--386 (2015)
+
+\bibitem{Toropov1989}
+Toropov, V.:
+\newblock Simulation approach to structural optimization.
+\newblock Structural Optimization \textbf{1}, 37--46 (1989)
+
+\bibitem{Toropov1992}
+Toropov, V.:
+\newblock Multipoint approximation method in optimization problems with expensive function values.
+\newblock In: Sydow, A. (ed.) Proceedings of the 4th international symposium on Systems analysis and simulation, 
+pp. 207--212, Elsevier,  (1992)
+
+\bibitem{ToropovFilatov1993}
+Toropov, V., Filatov, A., Polynkin, A.:
+\newblock Multiparameter structural optimization using fem and multipoint explicit approximations.
+\newblock Structural Optimization \textbf{6}, 7--14 (1993)
+
+\bibitem{ShahparPolynkinToropov2008}
+Shahpar, S., Polynkin, A., Toropov, V.:
+\newblock Large scale optimization of transonic axial compressor rotor blades.
+\newblock In:  49th AIAA/ASME/ASCE/AHS/ASC Structures, Structural Dynamics, and Materials Conference, AIAA, art. no. 2008-2056
+ (2008)
+
+\bibitem{PolynkinToropovShahpar2008}
+Polynkin, A., Toropov, V., Shahpar, S.:
+\newblock Design optimization of aircraft engine components.
+\newblock In: Proc. of 7th ASMO UK / ISSMO Conf. on Engineering Design Optimization. Process and Product Improvement (2008)
+
+\bibitem{PolynkinToropovShahpar2010}
+Polynkin, A., Toropov, V., Shahpar, S.:
+\newblock Multidisciplinary optimization of turbomachinery based on metamodel built by genetic programming.
+\newblock In: 13th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference (2010)
+
+\bibitem{KeulenToropovMarkine1996}
+van Keulen, F., Toropov, V., Markine, V.:
+\newblock Recent refinements in the multi-point approximation method in conjunction with adaptive mesh refinement.
+\newblock In: McCarthy, J.M. (ed.) Proc. of ASME Design Engineering Technical Conferences and Computers
+  in Engineering Conf., 1--12, ASME, Irvine CA (1996)
+
+\bibitem{ToropovKeulenMarkine1996}
+Toropov, V., van Keulen, F., Markine, V., de~Boer, H.:
+\newblock Refinements in the multi-point approximation method to reduce the effects of noisy responses.
+\newblock In: 6th AIAA/NASA/ISSMO Symp. on Multidisciplinary Analysis and Optimization,  pp. 941--951, AIAA (1996)
+
+\bibitem{ToropovMarkineHolden1999}
+Toropov, V., Markine, V., Holden, C.:
+\newblock Use of mid-range approximations for optimization problems with functions of domain-dependent calculability.
+\newblock In: 3rd ISSMO/UBCAD/UB/AIAA World Congress of Structural and Multidisciplinary Optimization (1999).
+
+\bibitem{ToropovMarkine1996}
+Toropov, V., Markine, V.:
+\newblock The use of simplified numerical models as mid-range approximations.
+\newblock In: 6th AIAA/NASA/ISSMO Symp. on Multidisciplinary Analysis and Optimization, pp. 952--958 (1996).
+
+\bibitem{ToropovAlvarez1998}
+Toropov, V., Alvarez, L.:
+\newblock Creation of multipoint approximations using genetic programming.
+\newblock In: Parmee, I.C. (ed.) Adaptive Computing in Design and Manufacture, 3rd Int. Conf.,
+   pp. 21--24, PEDC, Dartington (1998)
+
+\bibitem{PolynkinToropov2012}
+Polynkin, A. Toropov, V.:
+\newblock Mid-range metamodel assembly building based on linear regression for large scale optimization problems.
+\newblock Structural and Multidisciplinary Optimization \textbf{45}(4), 515--527 (2012)
+
+\bibitem{LancasterSalkauskas1981}
+Lancaster, P. Salkauskas, K.:
+\newblock Surfaces generated by moving least squares methods.
+\newblock Mathematics of Computation \textbf{87}, 141--158 (1981)
+
+\bibitem{Liszka1984}
+Liszka, T.:
+\newblock An interpolation method for an irregular net of nodes.
+\newblock Int. J. Num. Meth. \textbf{20}, 1599--1612 (1984)
+
+\bibitem{ChoiYounYang2001}
+Choi, K., Youn, B., Yang, R.-J.:
+\newblock Moving least squares method for reliability-based design optimization.
+\newblock In: 4th World Congress of Structural and Multidisciplinary Optimization, Dalian, China (2001)
+
+\bibitem{ToropovSchrammSahaiJones2005}
+Toropov, V., Schramm, U., Sahai, A., Jones, R., Zeguer, T.:
+\newblock Design optimization and stochastic analysis based on the moving least squares method.
+\newblock In: Herskovits, J., Mazorche, S., Canelas, A. (eds.) 6th World Congress of Structural and Multidisciplinary Optimization, art. no. 9412 (2005)
+
+\bibitem{PolynkinToropov2010}
+Polynkin, A. Toropov, V.:
+\newblock Recognition of design variable inter-dependencies using  cross-validated moving least-squares method.
+\newblock In: Proceedings of the 51st AIAA/ASME/ASCE/AHS/ASC Structures, Structural Dynamics, and Materials Conference, AIAA, art. no. 2010-2985 (2010)
+
+\bibitem{OllarToropovJones2014}
+Ollar, J., Toropov, V., Jones. R.:
+\newblock  Mid-range approximations in sub-spaces for MDO problems with disparate discipline attributes.
+\newblock In: 15th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference, AIAA, art. no. 2014-2437 (2014)
+
+\bibitem{KochYangGu2004}
+Koch, P.~N., Yang, R.-J., Gu, L.:
+\newblock Design for six sigma through robust optimization.
+\newblock Structural and Multidisciplinary Optimization \textbf{26}, 235--248 (2004)
+
+\bibitem{Yao2011}
+Yao, W.:
+\newblock Review of uncertainty-based multidisciplinary design optimization methods for aerospace vehicles.
+\newblock Progress in Aerospace Sciences \textbf {47}, 450--479 (2011)
+
+\bibitem{ZhangZhuChenArendt2013}
+Zhang, S., Zhu, P., Chen, W., Arendt, P.:
+\newblock Concurrent treatment of parametric uncertainty and metamodeling uncertainty in robust design.
+\newblock Structural and Multidisciplinary Optimization \textbf {47}(1), 63--76 (2013)
+
+\bibitem{AouesChateauneuf2010}
+Aoues, Y., Chateauneuf, A.:
+\newblock Benchmark study of numerical methods for reliability-based design optimization.
+\newblock Structural and Multidisciplinary Optimization \textbf {41}(2), 277--294 (2010)
+
+\bibitem{YounChoi2004}
+Youn, B.~D., Choi, K.~K.:
+\newblock Selecting probabilistic approaches for reliability-based design optimization.
+\newblock AIAA journal \textbf{42}(1), 124--131 (2004)
+
+\bibitem{RockafellarRoyset2010}
+Rockafellar, R., Royset, J.:
+\newblock On buffered failure probability in design and optimization of structures.
+\newblock Reliability Engineering \& System Safety \textbf{95}(5), 499--510 (2010)
+
+\bibitem{BasovaRockafellarRoyset}
+Basova, H., Rockafellar, R., Royset, J.:
+\newblock A computational study of buffered failure probability in reliability-based optimization.
+\newblock In: Proceedings of the 11th International Conference on Application of Statistics and Probability in Civil Engineering, Zurich, Switzerland (2011)
+
+\bibitem{RockafellarRoyset2015}
+Rockafellar, R.~T., Royset, J.:
+\newblock Engineering decisions under risk-averseness.
+\newblock ASCE-ASME Journal of Risk and Uncertainty in Engineering Systems, Part A: Civil Engineering
+ \textbf{1}(2), art. no. 04015003 (2015)
+
+\bibitem{Caflisch1998}
+Caflisch, R.:
+\newblock Monte Carlo and quasi-Monte Carlo methods.
+\newblock Acta Numerica \textbf{7}, 1--49 (1998)
+
+\bibitem{Niederreiter1992}
+Niederreiter, H.:
+\newblock Random Number Generation and Quasi-Monte Carlo Methods
+\newblock SIAM (1992)
+
+\bibitem{Vanderpllaats2001}
+Vanderplaats, G.:
+\newblock Multidiscipline Design Optimization. 
+\newblock Vanderplaats Research \& Development, Inc., Colorado Springs, (2001)
+
+\end{thebibliography}
 

### 2018-01-09 2018/Mixed-Integer/LION2018.tex

\documentclass{llncs}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage[misc,geometry]{ifsym}

\usepackage{amsmath}
\usepackage{enumitem}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage{epstopdf}
%\usepackage{epsfig}
\usepackage[T2A]{fontenc}
\usepackage[utf8x]{inputenc}
\usepackage[russian,english]{babel}

\begin{document}

\mainmatter

\title{Mixed-integer}
\author{Victor Gergel \and Konstantin Barkalov \and Ilya Lebedev %\Letter 
\\
\email{konstantin.barkalov@itmm.unn.ru}}

\institute{
Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
}

\maketitle

\begin{abstract}
The paper presents .

\keywords global optimization, non-convex constraints, mixed-integer problems.

\end{abstract}

\section{Introduction}\label{sec:intro}
\Russian
В работе рассматриваются mixed-integer global optimization problems и методы их решения. Задачи глобальной оптимизации обладают значительной вычислительной трудоемкостью, т.к. глобальный оптимум является интегральной характеристикой решаемой задачи и требует исследования всей области поиска. Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (вообще говоря, неравномерного) в области параметров. Особый интерес представляют задачи, в которых часть параметров может принимать лишь целочисленные значения, т.к. для них сложнее построить оценки оптимума по сравнению с непрерывными задачами.

Решению mixed-integer problems посвящена обширная литература (см., например, обзоры \cite{Burer,Boukouvala} и статьи, цитированные в них). Известные детерминированные методы решения задач данного класса основаны, как правило, на идеях Branch-and-Bound or Branch-and-Reduce. Также известен ряд метаэвристических и генетических алгоритмов, так или иначе основанных на идеях случайного поиска.

В данной работе нами предложен новый детерминированный метод решения mixed-integer problems, основанный на information-statistical approach к решению задач глобальной оптимизации, разработанном в ННГУ \cite{Strongin2000,Strongin2013}. Текст статьи, отражающей предварительные результаты исследования, построен следующим образом. Вначале рассмотрим подход к решению задач с непрерывными параметрами, затем будет предложено его обобщение для mixed-integer problems. В финальном разделе приведены результаты сравнения предложенного метода с известными аналогами.

\section{Global search algorithm and dimension reduction}

Let us consider a multiextremal optimization problem in the form
\begin{gather}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.
\end{gather}
Suppose, that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and
the left-hand sides $g_i(y), \; 1\leq i \leq m$, of the constraints satisfy the Lipschitz condition
\[ 
\left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in S, \; 1\leq i \leq m+1, 
\]
with constants $L_i, \; 1 \leq i \leq m+1$, respectively,  and may be multiextremal.

Using a continuous single-valued mapping $y(x)$  (Peano-type space-filling curve) of the interval $[0,1]$ onto $D$, a multidimensional problem (\ref{problem}) can be reduced to a one-dimensional problem
\begin{equation}\label{problem1}
\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
\end{equation}
The reduction of dimensionality matches the multidimensional problem with a Lipschitzian objective function and Lipschitzian constraints with a one-dimensional problem where the respective functions satisfy the uniform H\"older condition (see \cite{Strongin2000})
\[
\left|g_i(y(x'))-g_i (y(x''))\right| \leq H_i \left|x'-x'' \right|^{1/N}, \; x',x''\in [0,1], \; 1\leq i \leq m+1.
\]
Here $N$ is the dimensionality of the initial multidimensional problem and the coefficients $H_i$ are related with the Lipschitz constants $L_i$ of the initial problem by the inequalities $H_i \leq 2L_i \sqrt{N+3}$. The issues around the numeric construction of a Peano-type curve and the corresponding theory are considered in detail in \cite{Strongin2000,Strongin2013}.

An efficient global search algorithm (GSA) for solving one-dimensional constrained optimization problem (\ref{problem1}) is developed in UNN. \Russian Основная схема алгоритма состоит в следующем.

The first iteration $x^1 \in (0,1)$ is executed in an arbitrary internal point of the search interval. The choice of the next iteration point $x^{k+1}, \; k \geq 1,$ is performed according to the following rules.
\begin{itemize}
	\item Arrange the search information in the order of values of the iteration points executed earlier, i.e. 
	\[ 
	x_1 < x_2 <... <x_k.
	\]
	\item Compute a quantity $R(i)$ called a \textit{characteristic} of the interval for each interval $(x_{i-1},x_i), 1<i\leq k$.
	\item Determine the interval $(x_{t-1},x_t)$ which the maximum characteristic $R(t)=\max{\{R(i): 1\leq i \leq k}\}$ corresponds to.
	\item Compute the next iteration point $x^{k+1}$ in the interval with the maximum characteristic, i.e. $x^{k+1}\in (x_{t-1},x_t)$.
\end{itemize}

A detailed description of various variants of this algorithm is presented in \cite{ }.

\section{Global search algorithm for mixed-integer problems}

\Russian
Рассмотрим теперь способ адаптации GSA для решения mixed-integer global optimization problems вида

\begin{gather}\label{problem_i}
\min{\left\{ g_{m+1}(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
D=\left\{a_i\leq y_i \leq b_i, \; 1\leq i \leq N, \; y_j \in Z, \; j \in J, \; y_i \in R, \; i \notin J \right\}.
\end{gather}

На основе задачи (\ref{problem_i}) сформируем набор задач 
\begin{gather}\label{problem_is}
\min{\left\{ g_{m+1}^s(y):y\in D^s, \; g_i^s(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
D^s=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, \; i \notin J \right\}, \; s\in\{1,...,S\}, 
\end{gather}
каждая из которых соответствует исходной задаче (\ref{problem_i}) с фиксированным набором целочисленных параметров. Число задач $S$ будет соответствовать числу возможных комбинаций целочисленных параметров.

Используя схему редукции размерности с помощью развертки $y(x)$ из (\ref{problem1}) можно составить единую задачу
\begin{equation}\label{problem_is1}
\min \left\{g_{m+1}^s(y(x)): x \in [0,S], \; g_i^s(y^s(x)) \leq 0, \; 1 \leq i \leq m\right\},
\end{equation}
где отображение $y^s(x)$  формируется как на основе исходной развертки $y(x)$ следующим образом
\[
y^s(x)=y(x-s), \; x\in[s-1,s],\; s\in\{1,...,S\}
\]
При этом objective function and constraints будут иметь разрыв в точках $s∈{1,...,S-1}$, поэтому данные точки рассматриваются как выколотые, значение objective function and constraints в них не определено.

Для иллюстрации на рис. 1 изображены графики функций, соответствующих задаче с одним непрерывным и одним бинарным параметром.

Применяя к решению задачи (\ref{problem_is1}) алгоритм глобального поиска, мы найдем решение задачи (\ref{problem_i}). При этом основная часть испытаний будет проведена в $s$-й подзадаче, решение которой соответствует решению исходной задачи (\ref{problem_i}). В остальных подзадачах будет проведена лишь незначительная часть испытаний, т.к. решения данных подзадач являются локально-оптимальными по отношению к решению $s$-й подзадачи. Сказанное подтверждается рис. 1, где штрихами обозначены точки испытаний, выполненных при решении данной задачи.

Таким образом, мы сформировали Mixed Integer Global Search Algorithm (MIGSA). Можно показать, что условия сходимости данного алгоритма будут следовать из условий сходимости его прототипа (GSA).

\section{Results of experiments}

\Russian
Проведем сравнение предложенного метода MIGSA с genetic algorithm for solving mixed-integer optimization problems, реализованным в Matlab Global Optimization Toolbox. В табл. 1 приведено число испытаний, потребовавшееся для решения известных тестовых mixed-integer problems данными методами. Для обоих методов была использована одинаковая точность поиска решения $10^{-2}$. Результаты экспериментов показывают превосходство метода MIGSA как по числу итераций, так и по времени работы.

\begin{table}
	\caption{Сравнение эффективности методов MIGSA и GA}
	\label{tab:1}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
		Test problem & \multicolumn{2}{c}{ GA } & & \multicolumn{2}{c}{MIGSA} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & $k$ & $t$ & & $k$ & $t$  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		 Problem 2 \cite{Floudas}&	2081 &	0.199 & &	417 &	0.04 \\
		 Problem 3 \cite{Floudas}& 	3851 &	0.353 & & 3324 &	0.107 \\
		 Problem 6 \cite{Floudas}&	2081 &	0.165 & &	118 &	0.001 \\
		 Problem 1 \cite{Deep}   &	2081 &	0.154 & &	66 &	0.0007 \\
		 Problem 2 \cite{Deep}   &	2081 &	0.191 & &	57 &	0.0006 \\
		 Problem 7 \cite{Deep}   &	2321 &	0.284 & & 372	 &	0.017 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

Для демонстрации надежности метода MIGSA нами были решены четыре серии по 100 многоэкстремальных mixed-integer problems, построенных на основе модифицированных классов задач Simple и Hard, сгенерированных генератором GKLS [Sergeyev]. В задачах было 5 целочисленных и 3 непрерывных параметра. Точность поиска решения равнялась $10^{-2}$. Были решены все задачи из серии, при этом для решения задач на основе класса Simple потребовалось среднем 11988 trials, а для решения задач на основе класса Hard - 24750 trials.

\section*{Acknowledgment}
The authors are grateful for the support provided by the Russian Science Foundation, project No. 16-11-10150.

\begin{thebibliography}{10}

\bibitem{Burer}
Burer, S., Letchford, A.N.: Non-convex mixed-integer nonlinear programming: A survey. Surveys in Operations Research and Management Science \textbf{17}, 97--106 (2012)

\bibitem{Boukouvala}
Boukouvala, F., Misener, R., Floudas, C.A.: Global optimization advances in Mixed-Integer Nonlinear Programming, MINLP, and Constrained Derivative-Free Optimization, CDFO. European Journal of Operational Research \textbf{252}, 701--727 (2016)

\bibitem{Strongin2000}
Strongin, R.G., Sergeyev, Y.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000) %; DOI: 10.1007/978-1-4615-4677-1

\bibitem{Strongin2013}
Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013) %;  DOI: 10.1007/978-1-4614-8042-6

\bibitem{Floudas}
Floudas, C.A., Pardalos, P.M.:  Handbook of Test Problems in Local and Global Optimization. Springer (1999)  %; DOI: 10.1007/978-1-4757-3040-1

\bibitem{Deep}
Deep, K., Singh, K. P., Kansal, M.L., Mohan, C.: A real coded genetic algorithm for solving integer and mixed integer optimization problems. Appl. Math. Comput. \textbf{212}(2), 505--518 (2009)

\end{thebibliography}

\end{document}
______________________________________________________________________

### 2018-01-09 2018/LION_Mixed-Integer/LION2018.tex

--- 2018-01-09 2018/Mixed-Integer/LION2018.tex
@@
 
-Решению mixed-integer problems посвящена обширная литература (см., например, обзоры \cite{Burer,Boukouvala} и статьи, цитированные в них).
+Методам решения mixed-integer problems посвящена обширная литература (см., например, обзоры \cite{Burer,Boukouvala}).
 Известные детерминированные методы решения задач данного класса основаны, как правило, на идеях Branch-and-Bound or Branch-and-Reduce.
@@
 
-В данной работе нами предложен новый детерминированный метод решения mixed-integer problems, основанный на information-statistical approach к решению задач глобальной оптимизации, разработанном в ННГУ \cite{Strongin2000,Strongin2013}.
+В данной работе нами предложен новый детерминированный метод решения mixed-integer problems, основанный на information-statistical approach к решению задач глобальной оптимизации \cite{Strongin2000,Strongin2013}.
 Текст статьи, отражающей предварительные результаты исследования, построен следующим образом.
@@
 \end{equation}
-The reduction of dimensionality matches the multidimensional problem with a Lipschitzian objective function and Lipschitzian constraints with a one-dimensional problem where the respective functions satisfy the uniform H\"older condition (see \cite{Strongin2000})
-\[
-\left|g_i(y(x'))-g_i (y(x''))\right| \leq H_i \left|x'-x'' \right|^{1/N}, \; x',x''\in [0,1], \; 1\leq i \leq m+1.
-\]
-Here $N$ is the dimensionality of the initial multidimensional problem and the coefficients $H_i$ are related with the Lipschitz constants $L_i$ of the initial problem by the inequalities $H_i \leq 2L_i \sqrt{N+3}$.
-The issues around the numeric construction of a Peano-type curve and the corresponding theory are considered in detail in \cite{Strongin2000,Strongin2013}.
+The reduction of dimensionality matches the multidimensional problem with a Lipschitzian objective function and Lipschitzian constraints with a one-dimensional problem where the respective functions satisfy the uniform H\"older condition (see \cite{Strongin2000}).
 
-An efficient global search algorithm (GSA) for solving one-dimensional constrained optimization problem (\ref{problem1}) is developed in UNN. \Russian Основная схема алгоритма состоит в следующем.
-
-The first iteration $x^1 \in (0,1)$ is executed in an arbitrary internal point of the search interval.
-The choice of the next iteration point $x^{k+1}, \; k \geq 1,$ is performed according to the following rules.
-\begin{itemize}
-	\item Arrange the search information in the order of values of the iteration points executed earlier, i.e. 
-	\[ 
-	x_1 < x_2 <... <x_k.
-	\]
-	\item Compute a quantity $R(i)$ called a \textit{characteristic} of the interval for each interval $(x_{i-1},x_i), 1<i\leq k$.
-	\item Determine the interval $(x_{t-1},x_t)$ which the maximum characteristic $R(t)=\max{\{R(i): 1\leq i \leq k}\}$ corresponds to.
-	\item Compute the next iteration point $x^{k+1}$ in the interval with the maximum characteristic, i.e. $x^{k+1}\in (x_{t-1},x_t)$.
-\end{itemize}
-
+An efficient global search algorithm (GSA) for solving one-dimensional constrained optimization problem (\ref{problem1}) is considered in \cite{Strongin2000}.
 A detailed description of various variants of this algorithm is presented in \cite{ }.
@@
 
-На основе задачи (\ref{problem_i}) сформируем набор задач 
+На основе задачи (\ref{problem_i}) сформируем множество задач 
 \begin{gather}\label{problem_is}
@@
 
-Используя схему редукции размерности с помощью развертки $y(x)$ из (\ref{problem1}) можно составить единую задачу
+Используя схему редукции размерности с помощью развертки $y(x)$ на основе множества задач (\ref{ }) можно составить единую задачу
 \begin{equation}\label{problem_is1}
@@
 \end{equation}
-где отображение $y^s(x)$  формируется как на основе исходной развертки $y(x)$ следующим образом
+где отображение $y^s(x)$  формируется на основе исходной развертки $y(x)$ следующим образом
 \[
@@
 \]
-При этом objective function and constraints будут иметь разрыв в точках $s∈{1,...,S-1}$, поэтому данные точки рассматриваются как выколотые, значение objective function and constraints в них не определено.
+При этом функции $g_i^s(y^s(x))$ могут иметь разрывы в точках $s\in \{1,...,S-1\}$, поэтому данные точки рассматриваются как выколотые, значение objective function and constraints в них не определено.
 

### 2018-01-11 2018/LION_Mixed-Integer/LION2018.tex
Before translation

@@
 
-\title{Mixed-integer}
+\title{A Global Optimization Algorithm for Non-Convex Constrained Mixed-Integer Problems}
 \author{Victor Gergel \and Konstantin Barkalov \and Ilya Lebedev %\Letter 
@@
 \begin{abstract}
-The paper presents .
+\Russian
+В работе рассматриваются mixed-integer global optimization problems.
+Предложен новый детерминированный алгоритм для решения задач данного класса, основанный на information-statistical approach к решению задач непрерывной глобальной оптимизации.
+Проведено сравнение данного алгоритма с известными аналогами, показывающее эффективность разработанного подхода.
+Устойчивая работа алгоритма подтверждена также решением серии из нескольких сотен mixed-integer global optimization problems.
 
@@
 \Russian
-В работе рассматриваются mixed-integer global optimization problems и методы их решения.
+В работе рассматриваются global optimization problems и методы их решения.
 Задачи глобальной оптимизации обладают значительной вычислительной трудоемкостью, т.к.
@@
 Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (вообще говоря, неравномерного) в области параметров.
-Особый интерес представляют задачи, в которых часть параметров может принимать лишь целочисленные значения, т.к.
+Особый интерес представляют задачи, в которых часть параметров может принимать лишь целочисленные значения (mixed-integer global optimization problems), т.к.
 для них сложнее построить оценки оптимума по сравнению с непрерывными задачами.
@@
 Методам решения mixed-integer problems посвящена обширная литература (см., например, обзоры \cite{Burer,Boukouvala}).
-Известные детерминированные методы решения задач данного класса основаны, как правило, на идеях Branch-and-Bound or Branch-and-Reduce.
+Известные детерминированные методы решения задач данного класса основаны, как правило, на подходах Branch-and-Bound or Branch-and-Reduce.
 Также известен ряд метаэвристических и генетических алгоритмов, так или иначе основанных на идеях случайного поиска.
@@
 Текст статьи, отражающей предварительные результаты исследования, построен следующим образом.
-Вначале рассмотрим подход к решению задач с непрерывными параметрами, затем будет предложено его обобщение для mixed-integer problems.
+Вначале рассмотрен подход к решению задач с непрерывными параметрами, затем предложено его обобщение для mixed-integer problems.
 В финальном разделе приведены результаты сравнения предложенного метода с известными аналогами.
@@
 \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
-D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.
+D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.\nonumber
 \end{gather}
@@
 \[ 
-\left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in S, \; 1\leq i \leq m+1, 
+\left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq i \leq m+1, 
 \]
@@
 \end{equation}
-The reduction of dimensionality matches the multidimensional problem with a Lipschitzian objective function and Lipschitzian constraints with a one-dimensional problem where the respective functions satisfy the uniform H\"older condition (see \cite{Strongin2000}).
+The reduction of dimensionality matches the multidimensional problem with a Lipschitzian objective function and constraints with a one-dimensional problem where the respective functions satisfy the uniform H\"older condition (see \cite{Strongin2000}).
 
-An efficient global search algorithm (GSA) for solving one-dimensional constrained optimization problem (\ref{problem1}) is considered in \cite{Strongin2000}.
-A detailed description of various variants of this algorithm is presented in \cite{ }.
+%Проверить
+An efficient global search algorithm (GSA) for solving constrained optimization problem (\ref{problem1}) was developed at University of Nizhni Novgorod.
+A detailed description of the algorithm and the corresponding convergence theory are presented in \cite{Strongin2000}.
 
@@
 \Russian
-Рассмотрим теперь способ адаптации GSA для решения mixed-integer global optimization problems вида
+Рассмотрим теперь способ адаптации GSA для решения mixed-integer global optimization problems
 
@@
 \min{\left\{ g_{m+1}(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
-D=\left\{a_i\leq y_i \leq b_i, \; 1\leq i \leq N, \; y_j \in Z, \; j \in J, \; y_i \in R, \; i \notin J \right\}.
+D=\left\{a_i\leq y_i \leq b_i, \; 1\leq i \leq N, \; y_j \in Z, \; j \in J, \; y_i \in R, \; i \notin J \right\}.\nonumber
 \end{gather}
@@
 \min{\left\{ g_{m+1}^s(y):y\in D^s, \; g_i^s(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
-D^s=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, \; i \notin J \right\}, \; s\in\{1,...,S\}, 
+D^s=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, \; i \notin J \right\}, \; s\in\{1,...,S\},\nonumber 
 \end{gather}
@@
 
-Используя схему редукции размерности с помощью развертки $y(x)$ на основе множества задач (\ref{ }) можно составить единую задачу
+Используя схему редукции размерности с помощью развертки $y(x)$ на основе множества задач (\ref{problem_is}) можно составить единую задачу
 \begin{equation}\label{problem_is1}
@@
 \[
-y^s(x)=y(x-s), \; x\in[s-1,s],\; s\in\{1,...,S\}
+y^s(x)=y(x-s), \; x\in[s-1,s],\; s\in\{1,...,S\}.
 \]
-При этом функции $g_i^s(y^s(x))$ могут иметь разрывы в точках $s\in \{1,...,S-1\}$, поэтому данные точки рассматриваются как выколотые, значение objective function and constraints в них не определено.
+Функции $g_i^s(y^s(x))$ могут иметь разрывы в точках $s\in \{1,...,S-1\}$, поэтому данные точки рассматриваются как выколотые, значение objective function and constraints в них не определено.
 
-Для иллюстрации на рис. 1 изображены графики функций, соответствующих задаче с одним непрерывным и одним бинарным параметром.
+Для иллюстрации на Fig. \ref{fig:1}  изображены графики функций, соответствующих задаче с одним непрерывным и одним бинарным параметром.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.7\textwidth]{fig1.jpg}
+    \caption{Mixed-integer global optimization problem}
+    \label{fig:1}
+\end{figure}
 
@@
 решения данных подзадач являются локально-оптимальными по отношению к решению $s$-й подзадачи.
-Сказанное подтверждается рис.
-1, где штрихами обозначены точки испытаний, выполненных при решении данной задачи.
+Сказанное подтверждается Fig.
+\ref{fig:1}, где штрихами обозначены точки испытаний, выполненных при решении данной задачи.
 
-Таким образом, мы сформировали Mixed Integer Global Search Algorithm (MIGSA). Можно показать, что условия сходимости данного алгоритма будут следовать из условий сходимости его прототипа (GSA).
+Таким образом, мы сформировали Mixed-Integer Global Search Algorithm (MIGSA), основанный на сведении задачи MINLP путем к задаче NLP.
+Можно доказать, что условия сходимости данного алгоритма будут следовать из условий сходимости его прототипа (GSA).
 
@@
 Проведем сравнение предложенного метода MIGSA с genetic algorithm for solving mixed-integer optimization problems, реализованным в Matlab Global Optimization Toolbox.
-В табл.
-1 приведено число испытаний, потребовавшееся для решения известных тестовых mixed-integer problems данными методами.
+В Table \ref{tab:1} приведено число испытаний, потребовавшееся для решения известных тестовых mixed-integer problems данными методами.
 Для обоих методов была использована одинаковая точность поиска решения $10^{-2}$.
+Все вычислительные эксперименты были проведены на компьютере с процессором Intel Core i5-7300 2.5 GHz и 8 Gb RAM под управлением MS Windows 10.
 Результаты экспериментов показывают превосходство метода MIGSA как по числу итераций, так и по времени работы.
@@
 		\hline\noalign{\smallskip}
-		Test problem & \multicolumn{2}{c}{ GA } & & \multicolumn{2}{c}{MIGSA} \\
+	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c}{ GA } & & \multicolumn{2}{c}{MIGSA} \\
 		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
@@
 		\noalign{\smallskip} \hline \noalign{\smallskip}
-		 Problem 2 \cite{Floudas}&	2081 &	0.199 & &	417 &	0.04 \\
-		 Problem 3 \cite{Floudas}& 	3851 &	0.353 & & 3324 &	0.107 \\
-		 Problem 6 \cite{Floudas}&	2081 &	0.165 & &	118 &	0.001 \\
-		 Problem 1 \cite{Deep}   &	2081 &	0.154 & &	66 &	0.0007 \\
-		 Problem 2 \cite{Deep}   &	2081 &	0.191 & &	57 &	0.0006 \\
-		 Problem 7 \cite{Deep}   &	2321 &	0.284 & & 372	 &	0.017 \\
+		 Problem 2 \cite{Floudas}&	481 &	0.0601 & &	417 &	0.04 \\
+%		 Problem 3 \cite{Floudas}& 	1821 &	0.1130 & & 3324 &	0.107 \\
+		 Problem 6 \cite{Floudas}&	641 &	0.0510 & &	118 &	0.001 \\
+		 Problem 1 \cite{Deep}   &	481 &	0.1378 & &	66 &	0.0007 \\
+		 Problem 2 \cite{Deep}   &	481 &	0.0473 & &	57 &	0.0006 \\
+		 Problem 7 \cite{Deep}   &	841 &	0.0736 & & 372	 &	0.017 \\
 		\noalign{\smallskip}\hline
@@
 
-Для демонстрации надежности метода MIGSA нами были решены четыре серии по 100 многоэкстремальных mixed-integer problems, построенных на основе модифицированных классов задач Simple и Hard, сгенерированных генератором GKLS [Sergeyev].
-В задачах было 5 целочисленных и 3 непрерывных параметра.
-Точность поиска решения равнялась $10^{-2}$.
-Были решены все задачи из серии, при этом для решения задач на основе класса Simple потребовалось среднем 11988 trials, а для решения задач на основе класса Hard - 24750 trials.
+Для демонстрации надежности метода MIGSA нами были решены четыре серии по 100 многоэкстремальных mixed-integer problems, построенных на основе модифицированных классов задач Simple и Hard, сгенерированных генератором GKLS \cite{Gaviano}.
+Данный генератор многоэкстремальных функций часто используется для исследования алгоритмов непрерывной оптимизации \cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
+ В сгенерированных нами задачах было 5 целочисленных и 3 непрерывных параметра.
+Точность поиска решения равнялась $10^{-2}$ по координате.
+Все задачи из серии были успешно решены , при этом для решения задач на основе класса Simple потребовалось среднем 11988 trials, а для решения задач на основе класса Hard -- 24750 trials.
 
-\section*{Acknowledgment}
-The authors are grateful for the support provided by the Russian Science Foundation, project No. 16-11-10150.
+\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No 16-11-10150.
 
@@
 
+\bibitem{Gaviano}
+Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
+ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003)
+
+\bibitem{Paulavicius2014} 
+Paulavi\v{c}ius, R., Sergeyev, Y., Kvasov, D., \v{Z}ilinskas, J.: Globally-biased DISIMPL algorithm for expensive global optimization. J. Glob. Optim. \textbf{59}(2-3), 545--567 (2014)
+
+\bibitem{SergeyevKvasov2015} 
+Sergeyev, Y.D., Kvasov, D.E.: A deterministic global optimization using smooth diagonal auxiliary functions. Commun. Nonlinear. Sci. Numer. Simulat. \textbf{21}(1-3), 99--111 (2015)
+
+\bibitem{Lebedev2015}
+Lebedev, I., Gergel, V.: Heterogeneous Parallel Computations for Solving Global Optimization Problems. Procedia Computer Science \textbf{66}, 53--62 (2015)
+
+\bibitem{Gergel2015}
+Gergel, V., Sidorov, S.: A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems.
+Lecture Notes in Computer Science  \textbf{9251}, 505--515 (2015)
+
 \end{thebibliography}

### 2018-01-11 2018/LION_Mixed-Integer/LION2018.tex
After translation

@@
 %\usepackage{epsfig}
-\usepackage[T2A]{fontenc}
-\usepackage[utf8x]{inputenc}
-\usepackage[russian,english]{babel}
 
@@
 
-\title{A Global Optimization Algorithm for Non-Convex Constrained Mixed-Integer Problems}
+\title{A Global Optimization Algorithm for Non-Convex Mixed-Integer Problems}
 \author{Victor Gergel \and Konstantin Barkalov \and Ilya Lebedev %\Letter 
@@
 \begin{abstract}
-\Russian
-В работе рассматриваются mixed-integer global optimization problems.
-Предложен новый детерминированный алгоритм для решения задач данного класса, основанный на information-statistical approach к решению задач непрерывной глобальной оптимизации.
-Проведено сравнение данного алгоритма с известными аналогами, показывающее эффективность разработанного подхода.
-Устойчивая работа алгоритма подтверждена также решением серии из нескольких сотен mixed-integer global optimization problems.
+In the present paper, the mixed-integer global optimization problems are considered. A novel 
+deterministic algorithm for solving the problems of this class based on the information-statistical 
+approach to solving the continuous global optimization problems has been proposed. The 
+comparison of this algorithm with well known analogs demonstrating the efficiency of the 
+developed approach has been conducted. The stable operation of the algorithm was confirmed 
+also by solving a series of several hundred mixed-integer global optimization problems.
 
@@
 \section{Introduction}\label{sec:intro}
-\Russian
-В работе рассматриваются global optimization problems и методы их решения.
-Задачи глобальной оптимизации обладают значительной вычислительной трудоемкостью, т.к.
-глобальный оптимум является интегральной характеристикой решаемой задачи и требует исследования всей области поиска.
-Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (вообще говоря, неравномерного) в области параметров.
-Особый интерес представляют задачи, в которых часть параметров может принимать лишь целочисленные значения (mixed-integer global optimization problems), т.к.
-для них сложнее построить оценки оптимума по сравнению с непрерывными задачами.
 
-Методам решения mixed-integer problems посвящена обширная литература (см., например, обзоры \cite{Burer,Boukouvala}).
-Известные детерминированные методы решения задач данного класса основаны, как правило, на подходах Branch-and-Bound or Branch-and-Reduce.
-Также известен ряд метаэвристических и генетических алгоритмов, так или иначе основанных на идеях случайного поиска.
+In the present paper, the global optimization problems and the method of solving these ones are 
+considered. The global optimization problems are the computation-costly ones since the global 
+optimum is an integral characteristic of the problem being solved and requires the investigation 
+of the whole search domain. 
+%As a result, the search of the global optimum is reduced to the construction of a coverage (in general, a nonuniform one) of the space of parameters. 
+The problems, in which some parameters can take the inter values only (mixed-integer global 
+optimization problems) are of special interest because for these ones it is more difficult to build 
+the estimates of the optimum as compared to the continuous problems.
 
-В данной работе нами предложен новый детерминированный метод решения mixed-integer problems, основанный на information-statistical approach к решению задач глобальной оптимизации \cite{Strongin2000,Strongin2013}.
-Текст статьи, отражающей предварительные результаты исследования, построен следующим образом.
-Вначале рассмотрен подход к решению задач с непрерывными параметрами, затем предложено его обобщение для mixed-integer problems.
-В финальном разделе приведены результаты сравнения предложенного метода с известными аналогами.
+A lot of publication have been devoted to the methods of solving the mixed-integer problems 
+(see, for example, the reviews~\cite{Burer,Boukouvala}). The well known deterministic 
+methods of solving the problems of this class are based, as a rule, on the Branch-and-Bound 
+approach or on the Branch-and-Reduce one. Also, a number of the metaheuristic and genetic 
+algorithms are known based 
+ones, anyway, based on the ideas of random search.
+
+In the present study, we proposed a novel deterministic method for solving the mixed-integer 
+problems based on the information-statistical approach to solving the global optimization 
+problems \cite{Strongin2000,Strongin2013}. The paper text reflecting the results of preliminary 
+studies is composed as follows. First, an approach to solving the problems with continuous 
+parameters has been considered, then its generalization for the mixed-integer problems was 
+proposed. In the final section, the results of comparison of the proposed method with the well 
+known analogs are presented.
 
@@
 the left-hand sides $g_i(y), \; 1\leq i \leq m$, of the constraints satisfy the Lipschitz condition
-\[ 
-\left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq i \leq m+1, 
-\]
-with constants $L_i, \; 1 \leq i \leq m+1$, respectively,  and may be multiextremal.
+%\[ \left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq i \leq m+1, \]
+%with constants $L_i, \; 1 \leq i \leq m+1$, respectively,  
+and may be multiextremal.
 
@@
 
-%Проверить
-An efficient global search algorithm (GSA) for solving constrained optimization problem (\ref{problem1}) was developed at University of Nizhni Novgorod.
-A detailed description of the algorithm and the corresponding convergence theory are presented in \cite{Strongin2000}.
+An efficient global search algorithm (GSA) for solving the constrained optimization problem 
+(\ref{problem1}) has been developed at University of Nizhni Novgorod. A detailed description 
+of the algorithm and the corresponding convergence theory are presented in 
+\cite{Strongin2000}.
 
@@
 
-\Russian
-Рассмотрим теперь способ адаптации GSA для решения mixed-integer global optimization problems
-
+Now, let us consider a method of adaptation of GSA for solving the mixed-integer global 
+optimization problems 
 \begin{gather}\label{problem_i}
@@
 
-На основе задачи (\ref{problem_i}) сформируем множество задач 
+Let us formulate a set of problems based on the problem (\ref{problem_i})
 \begin{gather}\label{problem_is}
 \min{\left\{ g_{m+1}^s(y):y\in D^s, \; g_i^s(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
-D^s=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, \; i \notin J \right\}, \; s\in\{1,...,S\},\nonumber 
+D^s=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, \; i \notin J \right\}, \; s\in\{1,...,S\}.\nonumber 
 \end{gather}
-каждая из которых соответствует исходной задаче (\ref{problem_i}) с фиксированным набором целочисленных параметров.
-Число задач $S$ будет соответствовать числу возможных комбинаций целочисленных параметров.
+Each of these problems corresponds to the initial problem (\ref{problem_i}) with fixed set of 
+integer parameters. The number of problems $S$ will correspond to the number of possible 
+combinations of the integer parameters.
 
-Используя схему редукции размерности с помощью развертки $y(x)$ на основе множества задач (\ref{problem_is}) можно составить единую задачу
+Using the dimensionality reduction scheme utilizing the evolvent $y(x)$ 
+%based on the set of problems (\ref{problem_is}),
+one can compose a united problem 
 \begin{equation}\label{problem_is1}
-\min \left\{g_{m+1}^s(y(x)): x \in [0,S], \; g_i^s(y^s(x)) \leq 0, \; 1 \leq i \leq m\right\},
+\min \left\{g_{m+1}^s(y^s(x)): x \in [0,S], \; g_i^s(y^s(x)) \leq 0, \; 1 \leq i \leq m\right\},
 \end{equation}
-где отображение $y^s(x)$  формируется на основе исходной развертки $y(x)$ следующим образом
+where the mapping $y^s(x)$ is formed on the basis of initial evolvent $y(x)$ as follows
 \[
@@
 \]
-Функции $g_i^s(y^s(x))$ могут иметь разрывы в точках $s\in \{1,...,S-1\}$, поэтому данные точки рассматриваются как выколотые, значение objective function and constraints в них не определено.
+The functions $g_i^s(y^s(x))$ can have the breaks at the points $s\in \{1,...,S-1\}$, therefore, 
+these points were considered as the punctured ones, the values of the objective function and 
+constraints at these points remain undefined.
 
-Для иллюстрации на Fig. \ref{fig:1}  изображены графики функций, соответствующих задаче с одним непрерывным и одним бинарным параметром.
+As an illustration, in Fig. \ref{fig:1} presents the plots of the functions corresponding to a 
+problem with one continuous parameter and one binary parameter.
 
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{fig1.jpg}
+    \includegraphics[height=2.9cm,width=0.8\textwidth]{fig1.jpg}
     \caption{Mixed-integer global optimization problem}
@@
 
-Применяя к решению задачи (\ref{problem_is1}) алгоритм глобального поиска, мы найдем решение задачи (\ref{problem_i}).
-При этом основная часть испытаний будет проведена в $s$-й подзадаче, решение которой соответствует решению исходной задачи (\ref{problem_i}).
-В остальных подзадачах будет проведена лишь незначительная часть испытаний, т.к.
-решения данных подзадач являются локально-оптимальными по отношению к решению $s$-й подзадачи.
-Сказанное подтверждается Fig.
-\ref{fig:1}, где штрихами обозначены точки испытаний, выполненных при решении данной задачи.
+Applying the global search algorithm to solving the problem (\ref{problem_is1}), we will find 
+the solution of the problem (\ref{problem_i}). In this case the major part of trials will conducted in 
+the $s$-th subproblem, the solving of which corresponds to the solving of the initial 
+problem (\ref{problem_i}). In the rest subproblems, a minor part of trials only will be performed 
+since the solutions of these subproblems are the locally optimal ones with respect to the solution 
+of the $s$-th subproblem. All the above is confirmed by Fig. \ref{fig:1}, where the 
+points of trials executed in the course of solving this problem are denoted by the dashes.
 
-Таким образом, мы сформировали Mixed-Integer Global Search Algorithm (MIGSA), основанный на сведении задачи MINLP путем к задаче NLP.
-Можно доказать, что условия сходимости данного алгоритма будут следовать из условий сходимости его прототипа (GSA).
+Thus,  we have constructed the \textit{Mixed-Integer Global Search Algorithm} (MIGSA) based on the  
+reduction of the MINLP to the NLP. 
+%It is possible to prove that the convergence conditions for this algorithm will follow from the convergence conditions of its prototype (GSA).
 
@@
 
-\Russian
-Проведем сравнение предложенного метода MIGSA с genetic algorithm for solving mixed-integer optimization problems, реализованным в Matlab Global Optimization Toolbox.
-В Table \ref{tab:1} приведено число испытаний, потребовавшееся для решения известных тестовых mixed-integer problems данными методами.
-Для обоих методов была использована одинаковая точность поиска решения $10^{-2}$.
-Все вычислительные эксперименты были проведены на компьютере с процессором Intel Core i5-7300 2.5 GHz и 8 Gb RAM под управлением MS Windows 10.
-Результаты экспериментов показывают превосходство метода MIGSA как по числу итераций, так и по времени работы.
+Let us compare the proposed MIGSA method with a genetic algorithm for solving the mixed-
+integer optimization problems implemented in Matlab Global Optimization Toolbox. In Table 
+\ref{tab:1}, the numbers of trials required for solving the well-known test mixed-integer 
+problems by these methods are presented. For both methods, the same values of the accuracy of 
+the solution search $10^{-2}$ were used. All numerical experiments were conducted on a 
+computer with Intel Core i5-7300 2.5 GHz and 8 Gb RAM processor under MS Windows 10. The 
+results of experiments have demonstrated the advantage to the MIGSA method in the number 
+of iterations as well as in the execution time.
 
 \begin{table}
-	\caption{Сравнение эффективности методов MIGSA и GA}
+	\caption{Comparison of MIGSA and GA}
 	\label{tab:1}
@@
 
-Для демонстрации надежности метода MIGSA нами были решены четыре серии по 100 многоэкстремальных mixed-integer problems, построенных на основе модифицированных классов задач Simple и Hard, сгенерированных генератором GKLS \cite{Gaviano}.
-Данный генератор многоэкстремальных функций часто используется для исследования алгоритмов непрерывной оптимизации \cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
- В сгенерированных нами задачах было 5 целочисленных и 3 непрерывных параметра.
-Точность поиска решения равнялась $10^{-2}$ по координате.
-Все задачи из серии были успешно решены , при этом для решения задач на основе класса Simple потребовалось среднем 11988 trials, а для решения задач на основе класса Hard -- 24750 trials.
+In order to demonstrate the reliability of the MIGSA method, we have solved two series of 100 
+multiextremal mixed-integer problems each constructed on the basis of modified problems of 
+\textit{Simple} and \textit{Hard} classes generated by GKLS generator. %~\cite{Gaviano}. 
+This generator of multiextremal functions is often used for the investigations of the global optimization algorithms~ 
+\cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
+In the problems generated in our experiments, there were 5 integer parameters and 3 continuous 
+ones. The accuracy of the search %of the solution in the coordinates 
+was equal to $10^{-2}$. All  
+problems of the series have been solved successfully, at that, for solving the problems based on 
+the \textit{Simple} class 11988 trials in average were required whereas for solving the problems based 
+on the \textit{Hard} class 24750 trials were required.
 
@@
 Boukouvala, F., Misener, R., Floudas, C.A.: Global optimization advances in Mixed-Integer Nonlinear Programming, MINLP, and Constrained Derivative-Free Optimization, CDFO.
-European Journal of Operational Research \textbf{252}, 701--727 (2016)
+European J.
+Oper.
+Res.
+\textbf{252}, 701--727 (2016)
 
@@
 
-\bibitem{Gaviano}
-Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
+%\bibitem{Gaviano} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
 ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003)

### 2018-01-12 2018/LION_Mixed-Integer/LION2018.tex
After translation

@@
 approach to solving the continuous global optimization problems has been proposed. The 
-comparison of this algorithm with well known analogs demonstrating the efficiency of the 
+comparison of this algorithm with known analogs demonstrating the efficiency of the 
 developed approach has been conducted. The stable operation of the algorithm was confirmed 
@@
 %As a result, the search of the global optimum is reduced to the construction of a coverage (in general, a nonuniform one) of the space of parameters. 
-The problems, in which some parameters can take the inter values only (mixed-integer global 
+The problems, in which some parameters can take the integer values only (mixed-integer global 
 optimization problems) are of special interest because for these ones it is more difficult to build 
@@
 approach or on the Branch-and-Reduce one. Also, a number of the metaheuristic and genetic 
-algorithms are known based 
-ones, anyway, based on the ideas of random search.
+algorithms are known, which are based one way or another on the random search concept.
+%  based ones, anyway, based on the ideas of random search.
 
@@
 parameters has been considered, then its generalization for the mixed-integer problems was 
-proposed. In the final section, the results of comparison of the proposed method with the well 
+proposed. In the final section, the results of comparison of the proposed method with the 
 known analogs are presented.
@@
 
-Using a continuous single-valued mapping $y(x)$  (Peano-type space-filling curve) of the interval $[0,1]$ onto $D$, a multidimensional problem (\ref{problem}) can be reduced to a one-dimensional problem
+Using a continuous single-valued mapping $y(x)$  (Peano-type space-filling curve, or \textit{evolvent}) of the interval $[0,1]$ onto $D$, a multidimensional problem (\ref{problem}) can be reduced to a one-dimensional problem
 \begin{equation}\label{problem1}
-\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
+%\varphi(y(x^\ast))=
+\min \left\{g_{m+1}(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
 \end{equation}
@@
 \begin{gather}\label{problem_is}
-\min{\left\{ g_{m+1}^s(y):y\in D^s, \; g_i^s(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
-D^s=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, \; i \notin J \right\}, \; s\in\{1,...,S\}.\nonumber 
+\min{\left\{ g_{m+1}^s(y):y\in D, \; g_i^s(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
+D=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, \; i \notin J \right\}, \; s\in\{1,...,S\}.\nonumber 
 \end{gather}
-Each of these problems corresponds to the initial problem (\ref{problem_i}) with fixed set of 
-integer parameters. The number of problems $S$ will correspond to the number of possible 
+Each of these problems corresponds to the initial problem (\ref{problem_i}) with some combination of 
+integer parameters. The number of problems $S$ will correspond to the number of all possible 
 combinations of the integer parameters.
@@
     \includegraphics[height=2.9cm,width=0.8\textwidth]{fig1.jpg}
-    \caption{Mixed-integer global optimization problem}
+    \caption{Reduced mixed-integer global optimization problem}
     \label{fig:1}
@@
 the solution of the problem (\ref{problem_i}). In this case the major part of trials will conducted in 
-the $s$-th subproblem, the solving of which corresponds to the solving of the initial 
+the subproblem, the solving of which corresponds to the solving of the initial 
 problem (\ref{problem_i}). In the rest subproblems, a minor part of trials only will be performed 
-since the solutions of these subproblems are the locally optimal ones with respect to the solution 
-of the $s$-th subproblem. All the above is confirmed by Fig. \ref{fig:1}, where the 
+since the solutions of these subproblems are the locally optimal ones.
+%with respect to the solution of the $s$-th subproblem. 
+All the above is confirmed by Fig. \ref{fig:1}, where the  
 points of trials executed in the course of solving this problem are denoted by the dashes.
@@
 
-Let us compare the proposed MIGSA method with a genetic algorithm for solving the mixed-
-integer optimization problems implemented in Matlab Global Optimization Toolbox. In Table 
-\ref{tab:1}, the numbers of trials required for solving the well-known test mixed-integer 
-problems by these methods are presented. For both methods, the same values of the accuracy of 
-the solution search $10^{-2}$ were used. All numerical experiments were conducted on a 
-computer with Intel Core i5-7300 2.5 GHz and 8 Gb RAM processor under MS Windows 10. The 
+Let us compare the proposed MIGSA method with a genetic algorithm for solving the mixed-integer optimization problems implemented in Matlab Global Optimization Toolbox. In Table 
+\ref{tab:1}, the numbers of trials required for solving the known test mixed-integer 
+problems by these methods are presented. For both methods, the same accuracy of search $10^{-2}$ were used. All numerical experiments were conducted on a 
+computer with Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM under MS Windows 10. The 
 results of experiments have demonstrated the advantage to the MIGSA method in the number 
@@
 \cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
-In the problems generated in our experiments, there were 5 integer parameters and 3 continuous 
-ones. The accuracy of the search %of the solution in the coordinates 
+In the problems generated in our experiments, there were 5 integer  and 3 continuous parameters.
+The accuracy of the search %of the solution in the coordinates 
 was equal to $10^{-2}$. All  
-problems of the series have been solved successfully, at that, for solving the problems based on 
+problems of the series have been solved successfully and for solving the problems based on 
 the \textit{Simple} class 11988 trials in average were required whereas for solving the problems based 

### 2018-01-12 2018/LION_Toropov/LION2018.tex
Minor corrections

--- 2017-12-12 2018/LION/LION2018.tex
@@
 
-Optimization under uncertainty deals with responses $F(\pmb x,\pmb y)$ that depend on deterministic design variables $\pmb x$ and random ‘environmental’ variables $\pmb y$.
+Optimization under uncertainty deals with responses $F(\boldsymbol x,\boldsymbol y)$ that depend on deterministic design variables $\boldsymbol x$ and random ``environmental'' variables $\boldsymbol y$.
 Uncertainty in the design variables can also be modelled by introducing additional random variables, e.g.
 as additive or multiplicative noise.
-In this case, the responses $F(\pmb x,\pmb y)$ receive as input the sum $\pmb x+\pmb u$ (if the noise is additive), where $\pmb u$ is a random variable with a known probability distribution, yielding responses depending on $\pmb x,\; \pmb u$ and $\pmb y$.
-Randomness in $\pmb u$ and $\pmb y$ induces randomness in the responses $F(\pmb x+\pmb u,\pmb y)$.
+In this case, the responses $F(\boldsymbol x,\boldsymbol y)$ receive as input the sum $\boldsymbol x+\boldsymbol u$ (if the noise is additive), where $\boldsymbol u$ is a random variable with a known probability distribution, yielding responses depending on $\boldsymbol x,\; \boldsymbol u$ and $\boldsymbol y$.
+Randomness in $\boldsymbol u$ and $\boldsymbol y$ induces randomness in the responses $F(\boldsymbol x+\boldsymbol u,\boldsymbol y)$.
 To convert them into deterministic functions, that can be optimized, a mapping $R$, called the \textit{risk measure} \cite{RockafellarUryasev2000} or \textit{robustness measure} \cite{SchillingsSchulz2015}, is applied to the responses.
-The result is then a deterministic function of the design variables: $\widetilde{F}(\pmb x)=R(F(\pmb x+\pmb u, \pmb y))$.
+The result is then a deterministic function of the design variables: $\widetilde{F}(\boldsymbol x)=R(F(\boldsymbol x+\boldsymbol u, \boldsymbol y))$.
 The choice of a particular risk measure $R$ reflects the designer's attitude towards the risk of sub-optimal performance or constraint violation.
-For example, $R$ can be a combination of the mean and standard deviation of $F(\pmb x+ \pmb u,\pmb y)$ with respect to the joint probability distribution of $\pmb u$ and $\pmb y$ or a quantile of the distribution of $F(\pmb x+ \pmb u,\pmb y)$, i.e.
-such a value that the random variable of $F(\pmb x+\pmb u,\pmb y)$ exceeds it with a given, typically small, probability.
+For example, $R$ can be a combination of the mean and standard deviation of $F(\boldsymbol x+ \boldsymbol u,\boldsymbol y)$ with respect to the joint probability distribution of $\boldsymbol u$ and $\boldsymbol y$ or a quantile of the distribution of $F(\boldsymbol x+ \boldsymbol u,\boldsymbol y)$, i.e.
+such a value that the random variable of $F(\boldsymbol x+\boldsymbol u,\boldsymbol y)$ exceeds it with a given, typically small, probability.
 
-Computations of the risk measures typically involve evaluation of high-dimensional integrals, which becomes a challenging task if the number of design variables and/or environmental variables is large.
+Computations of the risk measures typically involve evaluation of high-di\-mensional integrals, which becomes a challenging task if the number of design variables and/or environmental variables is large.
 The problem becomes even more difficult if the original responses are computationally expensive, which is the case, for example, in computational fluid dynamics (CFD), where one function evaluation may take hours of even days.
@@
   \begin{array}{c}
-  \min\limits_{a_i \le x_i \le b_i}F_0(\pmb x) \\
-  s.t.\; F_j(\pmb x) \le 1,\; j=1,\dots ,M,
+  \min\limits_{a_i \le x_i \le b_i}F_0(\boldsymbol x) \\
+  s.t.\; F_j(\boldsymbol x) \le 1,\; j=1,\dots ,M,
   \end{array}
 \end{equation}
-where $\pmb x$ is the vector of design variables, $\pmb a$ and $\pmb b$ are the lower and upper bounds for the design variables, respectively, $F_0(\pmb x)$ is the objective function, and $F_j(\pmb x)$ are the constraints.
+where $\boldsymbol x$ is the vector of design variables, $\boldsymbol a$ and $\boldsymbol b$ are the lower and upper bounds for the design variables, respectively, $F_0(\boldsymbol x)$ is the objective function, and $F_j(\boldsymbol x)$ are the constraints.
 The numbers of design variables and constraints are $n$ and $M$, respectively.
@@
 \begin{enumerate}
-\item Initialization: choose a starting point $\pmb x^0$ and initial trust region $[\pmb a^0, \pmb b^0]$ such that $\pmb x^0 \in [\pmb a^0, \pmb b^0]$.
-\item At the $k$-th iteration the current approximation to the constrained minimum is $\pmb x^k$, the current trust region is $[\pmb a^k, \pmb b^k] \subset [\pmb a^0, \pmb b^0]$.
+\item Initialization: choose a starting point $\boldsymbol x^0$ and initial trust region $[\boldsymbol a^0, \boldsymbol b^0]$ such that $\boldsymbol x^0 \in [\boldsymbol a^0, \boldsymbol b^0]$.
+\item At the $k$-th iteration the current approximation to the constrained minimum is $\boldsymbol x^k$, the current trust region is $[\boldsymbol a^k, \boldsymbol b^k] \subset [\boldsymbol a^0, \boldsymbol b^0]$.
   \begin{enumerate}[label=(\alph*)]
-    \item Design of Experiments (DoE): a set of points $\pmb x_k^i \in [\pmb a^k, \pmb b^k]$ is chosen to be used for approximation building.
+    \item Design of Experiments (DoE): a set of points $\boldsymbol x_k^i \in [\boldsymbol a^k, \boldsymbol b^k]$ is chosen to be used for approximation building.
 Responses are evaluated at the DoE points and approximations are built using the obtained values.
@@
 
-    Denote the approximate objective function and constraints by $\widetilde{F}_0(\pmb x)$ and $\widetilde{F}_j(\pmb x)$, respectively.
+    Denote the approximate objective function and constraints by $\widetilde{F}_0(\boldsymbol x)$ and $\widetilde{F}_j(\boldsymbol x)$, respectively.
     \item The original optimization problem (\ref{eq:problem}) is replaced by the following problem:
@@
       \begin{array}{c}
-      \min\limits_{a_i^k \le x_i ^k\le b_i}\widetilde{F}_0(\pmb x) \\
-      s.t.\; \widetilde{F}_j(\pmb x) \le 1,\; j=1,\dots ,M,
+      \min\limits_{a_i^k \le x_i\le b_i^k}\widetilde{F}_0(\boldsymbol x) \\
+      s.t.\; \widetilde{F}_j(\boldsymbol x) \le 1,\; j=1,\dots ,M.
       \end{array}
@@
     The approximate problem (\ref{eq:problem_approx}) is solved using Sequential Quadratic Programming (SQP) and the solution of this problem determines the center of the next trust region.
-    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region \cite{KeulenToropovMarkine1996}.
-    \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $\pmb x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations).
+    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $\boldsymbol x^k$, and on the size of the current trust region \cite{KeulenToropovMarkine1996}.
+    \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $\boldsymbol x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations).
 If the termination criterion is satisfied, the algorithm proceeds to step 3.
@@
   \end{enumerate}
-  \item Optimization terminates. The obtained approximation to the solution of the problem (\ref{eq:problem}) is $\pmb x^{k+1}$.
+  \item Optimization terminates. The obtained approximation to the solution of the problem (\ref{eq:problem}) is $\boldsymbol x^{k+1}$.
 \end{enumerate}
@@
   \begin{array}{c}
-  \min\limits_{a_i \le x_i \le b_i}F_0(\pmb x + \pmb u, \pmb y) \\
-  s.t.\; F_j(\pmb x + \pmb u, \pmb y) \le 1,\; j=1,\dots ,M,
+  \min\limits_{a_i \le x_i \le b_i}F_0(\boldsymbol x + \boldsymbol u, \boldsymbol y) \\
+  s.t.\; F_j(\boldsymbol x + \boldsymbol u, \boldsymbol y) \le 1,\; j=1,\dots ,M,
   \end{array}
 \end{equation}
-where the random variable $\pmb u$ represents (additive) noise in the design variables and $\pmb y$ is the environmental variable, representing input to the responses that cannot be influenced by the designer.
-This problem statement is not precise since the random responses $F_0 (\pmb x+\pmb u,\pmb y),\; F_j (\pmb x+\pmb u,\pmb y)$ cannot be directly optimized.
+where the random variable $\boldsymbol u$ represents (additive) noise in the design variables and $\boldsymbol y$ is the environmental variable, representing input to the responses that cannot be influenced by the designer.
+This problem statement is not precise since the random responses $F_0 (\boldsymbol x+\boldsymbol u,\boldsymbol y),\; F_j (\boldsymbol x+\boldsymbol u,\boldsymbol y)$ cannot be directly optimized.
 Therefore, a risk measure $R$ needs to be applied to the objective and the constraints:
@@
   \begin{array}{c}
-  \min\limits_{a_i \le x_i \le b_i}R_{\pmb u,\pmb y}(F_0)(\pmb x) \\
-  s.t.\; R_{\pmb u,\pmb y}(F_j)(\pmb x) \le 1,\; j=1,\dots ,M,
+  \min\limits_{a_i \le x_i \le b_i}R_{u,y}(F_0)(\boldsymbol x) \\
+  s.t.\; R_{u,y}(F_j)(\boldsymbol x) \le 1,\; j=1,\dots ,M,
   \end{array}
 \end{equation}
-where the subscript in $R_{\pmb u,\pmb y}$ indicates that the variables $\pmb u$ and $\pmb y$ ‘collapse’ when the risk measure is applied and the result only depends on $\pmb x$.
+where the subscript in $R_{u,y}$ indicates that the variables $\boldsymbol u$ and $\boldsymbol y$ ‘collapse’ when the risk measure is applied and the result only depends on $\boldsymbol x$.
 The choices of risk measure will be discussed in Section \ref{sec:risk}.
@@
 
-First, it needs to be decided on whether to build metamodels for the risk measure $R_{\pmb u,\pmb y}(F_j)(\pmb x)$ or for the original responses $F_j (\pmb x+ \pmb u,\pmb y)$.
+First, it needs to be decided on whether to build metamodels for the risk measure $R_{u,y}(F_j)(\boldsymbol x)$ or for the original responses $F_j (\boldsymbol x+ \boldsymbol u,\boldsymbol y)$.
 The first option may seem attractive due to the simplicity of its implementation in Algorithm 1 but the deciding argument has to be that of the computational cost associated with the evaluation of either function.
-On the one hand, one evaluation of the function $R_{\pmb u,\pmb y}(F_j)(\pmb x)$ involves computing one or several integrals of the original response $F_j(\pmb x+ \pmb u,\pmb y)$ over the distribution of $\pmb u$ and $\pmb y$, which requires many evaluations of the original response.
-On the other hand, the dimensionality of the argument of $R_{\pmb u,\pmb y}(F_j)(\pmb x)$ is less than that of the argument of $F_j(\pmb x+ \pmb u,\pmb y)$, therefore, in principle, fewer function evaluations are needed to approximate the former than the latter.
+On the one hand, one evaluation of the function $R_{u,y}(F_j)(\boldsymbol x)$ involves computing one or several integrals of the original response $F_j(\boldsymbol x+ \boldsymbol u,\boldsymbol y)$ over the distribution of $\boldsymbol u$ and $\boldsymbol y$, which requires many evaluations of the original response.
+On the other hand, the dimensionality of the argument of $R_{u,y}(F_j)(\boldsymbol x)$ is less than that of the argument of $F_j(\boldsymbol x+ \boldsymbol u,\boldsymbol y)$, therefore, in principle, fewer function evaluations are needed to approximate the former than the latter.
 In practice, this may be a matter of a trade-off between the two abovementioned effects that depends on the proportion between the numbers of design variables $n$ and the environmental variables $n_e$ and on the complexity of the integration method and metamodeling method used.
-If the number if design variables is significantly larger than the number of environmental variables, which is often the case and which is assumed in this paper, building metamodels for the original responses $F_j(\pmb x+ \pmb u,\pmb y)$ is preferred.
+If the number if design variables is significantly larger than the number of environmental variables, which is often the case and which is assumed in this paper, building metamodels for the original responses $F_j(\boldsymbol x+ \boldsymbol u,\boldsymbol y)$ is preferred.
 
@@
 \begin{enumerate}
-%  \item Initialization: choose a starting point $\pmb x_0 \in {\Bbb R}^n$. 
-  \item Initialization: choose a starting point $\pmb x_0 \in {\pmb R}^n$. 
-	Choose an initial trust region $[\pmb a^0,\pmb b^0]$ in the combined space of design variables and environmental variables. 
-	Ensure that the initial guess $\pmb x_0$ belongs to the projection of the initial trust region onto the subspace of design variables.
+%  \item Initialization: choose a starting point $\boldsymbol x_0 \in {\Bbb R}^n$. 
+  \item Initialization: choose a starting point $\boldsymbol x_0 \in {\boldsymbol R}^n$. 
+	Choose an initial trust region $[\boldsymbol a^0,\boldsymbol b^0]$ in the combined space of design variables and environmental variables. 
+	Ensure that the initial guess $\boldsymbol x_0$ belongs to the projection of the initial trust region onto the subspace of design variables.
 The projection of the initial trust region onto the subspace of environmental variables must cover the region where the realizations of the random variable y are likely to be.
@@
 Typically, the trust region along the coordinates corresponding to the $i$-th environmental variable is centered at the mean of the $i$-th environmental variable and has the width of six of its standard deviations.
-  \item At the $k$-th iteration the current approximation to the constrained minimum is $\pmb x^k$, the current trust region is $[\pmb a^k,\pmb b^k]$.
+  \item At the $k$-th iteration the current approximation to the constrained minimum is $\boldsymbol x^k$, the current trust region is $[\boldsymbol a^k,\boldsymbol b^k]$.
   \begin{enumerate}[label=(\alph*)]
-    \item Design of Experiments (DoE): a set of points $(x_i^k,y_i^k)\in [\pmb a^k,\pmb b^k]$ is chosen that will be used to build approximations.
+    \item Design of Experiments (DoE): a set of points $(x_i^k,y_i^k)\in [\boldsymbol a^k,\boldsymbol b^k]$ is chosen that will be used to build approximations.
 Responses are evaluated at the DoE points and approximations are built using the obtained values.
-    Denote the approximate objective function and constraints by $\widetilde{F}_0(\pmb x, \pmb y)$ and $\widetilde{F}_j(\pmb x, \pmb y)$, respectively.
+    Denote the approximate objective function and constraints by $\widetilde{F}_0(\boldsymbol x, \boldsymbol y)$ and $\widetilde{F}_j(\boldsymbol x, \boldsymbol y)$, respectively.
     \item The original optimization problem is replaced by the following problem:
@@
       \begin{array}{c}
-      \min\limits_{a_i \le x_i \le b_i}R_{\pmb u,\pmb y}(\widetilde{F}_0)(\pmb x) \\
-      s.t.\; R_{\pmb u,\pmb y}(\widetilde{F}_j)(\pmb x) \le 1,\; j=1,\dots ,M,
+      \min\limits_{a_i^k \le x_i \le b_i^k}R_{u,y}(\widetilde{F}_0)(\boldsymbol x) \\
+      s.t.\; R_{u,y}(\widetilde{F}_j)(\boldsymbol x) \le 1,\; j=1,\dots ,M,
       \end{array}
@@
     The risk measures are calculated using the metamodels obtained in the previous step.
-Integrals that arise during this process are calculated with respect to the joint distribution of $\pmb u$ and $\pmb y$.
+Integrals that arise during this process are calculated with respect to the joint distribution of $\boldsymbol u$ and $\boldsymbol y$.
 This distribution is assumed to be fixed.
-    The approximate problem (\ref{eq:problem_u_tilde}) is solved using Sequential Quadratic Programming (SQP) and the center of the next trust region $\pmb x^{k+1}$ is determined as the solution of this problem.
+    The approximate problem (\ref{eq:problem_u_tilde}) is solved using Sequential Quadratic Programming (SQP) and the center of the next trust region $\boldsymbol x^{k+1}$ is determined as the solution of this problem.
 The optimization runs in a subspace of the space where metamodels are built: metamodels are defined in the combined space of design variables and environmental variables, whilst optimization runs in the subspace of the design variables.
@@
 The size of the trust region along the dimensions corresponding to the environmental variables is unchanged.
-The size of the next trust region along the dimensions corresponding to the design variables is determined depending on the quality of approximations at the previous iteration, on the history of the points $\pmb x^k$, and on the size of the current trust region according to the trust region strategy explained in \cite{KeulenToropovMarkine1996}.
+The size of the next trust region along the dimensions corresponding to the design variables is determined depending on the quality of approximations at the previous iteration, on the history of the points $\boldsymbol x^k$, and on the size of the current trust region according to the trust region strategy explained in \cite{KeulenToropovMarkine1996}.
     \item The termination criterion is checked. If the termination criterion is satisfied, the algorithm proceeds to step 3. Otherwise, it returns to step 2
   \end{enumerate}
-  \item Optimization terminates. The obtained approximation to the optimum is $\pmb x^{k+1}$.
+  \item Optimization terminates. The obtained approximation to the optimum is $\boldsymbol x^{k+1}$.
 \end{enumerate}
@@
 
-As it has been already said, a risk measure is a mapping that transforms a random response $F(\pmb x+ \pmb u,\pmb y)$  into a deterministic function that can be optimized.
+As it has been already said, a risk measure is a mapping that transforms a random response $F(\boldsymbol x+ \boldsymbol u,\boldsymbol y)$  into a deterministic function that can be optimized.
 Next we provide some examples of risk measures \cite{RockafellarUryasev2000,SchillingsSchulz2015}.
@@
   \item $R(F)=q_F(\alpha)$, where $q_F(\alpha)$ is the value of the quantile function, or simply the quantile, corresponding to a given probability $\alpha\in (0,1)$.
-This choice corresponds to the paradigm of reliability-based design optimization  \cite{AouesChateauneuf2010,Yao2011,YounChoi2004}.
+This choice corresponds to the paradigm of \textit{reliability-based design optimization} \cite{AouesChateauneuf2010,Yao2011,YounChoi2004}.
 In this approach, one wants to keep the probability of the system failing (i.e.
@@
 Superquantiles exhibit significantly better numerical properties than usual quantiles \cite{RockafellarRoyset2010}.
-The $\alpha$-superquantile can also be interpreted either as an average of quantiles  for $\alpha < \beta <1$, or as the conditional expectation of $F$ in its $(1−\alpha)$-tail \cite{RockafellarRoyset2015}:
+The $\alpha$-superquantile can also be interpreted either as an average of quantiles  for $\alpha < \beta <1$, or as the conditional expectation of $F$ in its $(1-\alpha)$-tail \cite{RockafellarRoyset2015}:
 \[
@@
 To compute the safety margin risk measure, one needs to compute two integrals:
-\begin{displaymath}
-  \begin{array}{c}
-    \mu (F)=E[F]=\int F(\pmb x + \pmb u, \pmb y)p df(\pmb u, \pmb y)d\pmb u d\pmb y \\
-    \sigma^2(F)=E[(F-E[F])^2]=\\\int(F(\pmb x + \pmb u, \pmb y)-E[F])^2 p df(\pmb u, \pmb y)d\pmb u d\pmb y
-  \end{array}
-\end{displaymath}
+\begin{gather*}
+    \mu (F)=E[F]=\int F(\boldsymbol x + \boldsymbol u, \boldsymbol y)p df(\boldsymbol u, \boldsymbol y)d\boldsymbol u d\boldsymbol y \\
+    \sigma^2(F)=E[(F-E[F])^2]=\int(F(\boldsymbol x + \boldsymbol u, \boldsymbol y)-E[F])^2 p df(\boldsymbol u, \boldsymbol y)d\boldsymbol u d\boldsymbol y
+\end{gather*}
 
@@
     \begin{equation} \label{eq:auxiliary_problem}
-			\overline{q}_F(\alpha) = \min_c\{c+E[(F-c)_+]\},
+			\overline{q}_F(\alpha) = \min_c\{c+E\left[(F-c)_+\right]\},
     \end{equation}
-where where ($F-c)_+$ denotes the positive part of the function $(F(\pmb x+ \pmb u,\pmb y) - c)_+$ (i.e.
-$F(\pmb x+ \pmb u,\pmb y) - c)$, if this value is positive, and zero otherwise), $E[\cdot]$ is, as before, the expected value, and $c$ is a scalar.
+where where ($F-c)_+$ denotes the positive part of the function $F(\boldsymbol x+ \boldsymbol u,\boldsymbol y) - c$ (i.e.
+$F(\boldsymbol x+ \boldsymbol u,\boldsymbol y) - c$, if this value is positive, and zero otherwise), $E[\cdot]$ is, as before, the expected value, and $c$ is a scalar.
 It has been shown by Rockafellar and Uryasev \cite{RockafellarUryasev2000} that the function to be minimized in (\ref{eq:auxiliary_problem}) is convex and continuously differentiable in $c$ if the underlying distribution is continuous.
@@
 \end{displaymath}
-The moment of inertia of a segment is $I_i=\frac{b_i h_i^3}{12}$ and the bending moment at the left end is $M_i=P[L+l_i- \sum_{j=1}^{i}  l_j ]$.
+The moment of inertia of a segment is $I_i=b_i h_i^3 / 12$ and the bending moment at the left end is $M_i=P[L+l_i- \sum_{j=1}^{i}  l_j ]$.
 The maximum bending stress in the segment $i$ is then given by the following formula:
 \begin{displaymath}
-  \sigma_i=\frac{M_i h_i}{2I_i}
+  \sigma_i=\frac{M_i h_i}{2I_i}.
 \end{displaymath}
@@
   \begin{array}{c}
-    \min\limits_{\pmb b, \pmb h}V(\pmb b, \pmb h) \\
+    \min\limits_{\boldsymbol b, \boldsymbol h}V(\boldsymbol b, \boldsymbol h) \\
     s.t.\;1.0\le b_i \le 10.0, \\
     5.0 \le h_i \le 100.0, \\
-    y_N\le 2/5, \\
+    y_N\le 2.5, \\
     \sigma_i \le \bar{\sigma}=14000, \\
@@
   \begin{array}{c}
-    \min\limits_{\pmb b, \pmb h}\mu(V(\pmb b, \pmb h) + k\sigma V(\pmb b, \pmb h)) \\
+    \min\limits_{\boldsymbol b, \boldsymbol h}\mu(V(\boldsymbol b, \boldsymbol h) + k\sigma V(\boldsymbol b, \boldsymbol h)) \\
     s.t.\;1.0\le b_i \le 10.0, \\
@@
   \begin{array}{c}
-    \min\limits_{\pmb b, \pmb h} \overline{q}_\alpha \left(V(\pmb b, \pmb h)\right) \\
+    \min\limits_{\boldsymbol b, \boldsymbol h} \overline{q}_\alpha \left(V(\boldsymbol b, \boldsymbol h)\right) \\
     s.t.\;1.0 \le b_i \le 10.0, \\
@@
 The $\mu(F ) + k\sigma(F)$ risk measure uses quantities that are estimated in the middle of the distribution --- the mean and the standard deviation.
-On the one hand, this has the advantage that samples drawn directly from the distribution of the random variables $\pmb u$ and $\pmb y$ are well suited to approximately compute this risk measure.
+On the one hand, this has the advantage that samples drawn directly from the distribution of the random variables $\boldsymbol u$ and $\boldsymbol y$ are well suited to approximately compute this risk measure.
 On the other hand, probabilistic interpretation of the obtained results in this case is only possible under the normality assumption for the distribution of $F$; if the actual distribution of $F$ has heavier tails than the normal one, the actual probability of failure will be higher than the one  the system was designed for.

### 2018-01-17 2017/KAZAN_LJM/paper_lobach.tex

--- 2017-11-20 2017/paper_lobach.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%% file template-ljm.tex %%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% This is a general template file for the LaTeX package ljm-auth
-% for Lobachevskii Journal of Mathematics 2009/07/20
-%
-% Copy it to a new file with a new name and use it as the basis
-% for your article. Delete % signs as needed.
-%
+\documentclass[
+11pt,%
+tightenlines,%
+twoside,%
+onecolumn,%
+nofloats,%
+nobibnotes,%
+nofootinbib,%
+superscriptaddress,%
+noshowpacs,%
+centertags]%
+{revtex4}
+\usepackage{ljm}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[12pt]{amsart}
-
-\usepackage{ljm-auth}
-
-\usepackage{graphicx}
-
-\usepackage{cite}
-
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{corollary}{Corollary}
-\newtheorem{lemma}{Lemma}
-
-\author{Strongin R.G., Gergel V.P., Barkalov K.A., Sysoyev A.V.}
-\crauthor{Strongin, Gergel, Barkalov, Sysoyev} % for running heads
-
-\tit{Generalized Parallel Computational Schemes for Time-Consuming Global Optimization}
-\shorttit{Parallel Schemes for Global Optimization} %for running heads
-
-\setcounter{page}{1}
+\setcounter{page}{3}
 
 \begin{document}
+\titlerunning{Generalized Parallel Schemes for Global Optimization}
+\authorrunning{Strongin, Gergel, Barkalov, Sysoyev}
 
-\maketit
+\title{Generalized Parallel Computational Schemes\\
+for Time-Consuming Global Optimization}
 
-\address{Lobachevsky State University of Nizhni Novgorod}
+\author{\firstname{R.~G.}~\surname{Strongin}}
+\email[E-mail: ]{president@unn.ru}
+\affiliation{Lobachevsky State University of Nizhni Novgorod}
 
-\email{gergel@unn.ru}
+\author{\firstname{V.~P.}~\surname{Gergel}}
+\email[E-mail: ]{gergel@unn.ru}
+\affiliation{Lobachevsky State University of Nizhni Novgorod}
 
-\abstract{This paper addresses computationally intensive global optimization problems, for solving of which the supercomputing systems with exaflops performance can be required.
+\author{\firstname{K.~A.}~\surname{Barkalov}}
+\email[E-mail: ]{barkalov@vmk.unn.ru}
+\affiliation{Lobachevsky State University of Nizhni Novgorod}
+
+\author{\firstname{A.~V.}~\surname{Sysoyev}}
+\email[E-mail: ]{sysoyev@vmk.unn.ru}
+\affiliation{Lobachevsky State University of Nizhni Novgorod}
+
+\firstcollaboration{(Submitted by ) }
+
+\received{}
+
+\begin{abstract}
+This paper addresses computationally intensive global optimization problems, for solving of which the supercomputing systems with exaflops performance can be required.
 To overcome such computational complexity, the paper proposes the generalized parallel computational schemes, which may involve numerous efficient parallel algorithms of global optimization.
-The proposed schemes include various ways of multilevel decomposition of parallel computations to guarantee the computational efficiency of supercomputing systems with shared and distributed memory multiprocessors with thousands of processors to meet global optimization challenges.}
+The proposed schemes include various ways of multilevel decomposition of parallel computations to guarantee the computational efficiency of supercomputing systems with shared and distributed memory multiprocessors with thousands of processors to meet global optimization challenges.
+\end{abstract}
 
-\notes{0}
-{
-%\subclass{} % 2010 Mathematical Subject Classification
+\subclass{90C26, 90C30}
 
-\keywords{
-global optimization, information-statistical theory, parallel computations, high-performance computing, supercomputing technologies
-}
+\keywords{global optimization, information-statistical theory, parallel computations, high-performance computing, supercomputing technologies}
 
-\thank{
-This research was supported by the Russian Science Foundation, project No 16-11-10150 "Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance"
-}
-}
+\maketitle
 
@@
 In the context of such problems, the performance criteria are nonlinear mostly, the search domains may be disconnected, and, above all, computing the function values may be computational intensive.
+
+From all the above, it follows that solving the global optimization problems within reasonable times is possible only utilizing the exascale supercomputers, which use the high-performance parallel algorithms of global optimization
 
@@
 
+The numerical solution of a global optimization problem consists in finding a numerical estimate $\bar{y}^\ast$ of the global minimum point $y^\ast \in D$ with a required accuracy $\varepsilon > 0$, i.e.
+
+\[
+\|\bar{y}^\ast - y^\ast\| \leq \varepsilon.
+\]
+
 It is supposed also that the minimized function $\varphi(y)$ is defined by a computational procedure for computing the value $\varphi(y)$ at any point $y \in D$ (hereafter such computations will be called \textit{trials}).
@@
 This results in a very high computational intensity of global optimization problems even in the case of small dimensionality.
+For example, if a uniform grid is used to cover the search domain, the number of grid nodes increases exponentially with increasing problem dimensionality, i.e.
+\[
+K = K(\varepsilon) \approx \prod_{i=1}^N[(b_i - a_i) / \varepsilon],
+\]
+where $\varepsilon > 0$ is the uniform grid step (i.e. even for a grid with $10$ nodes per axis and $N = 20$, the total number of nodes is as large as $10^{20}$).
 
 The computational complexity can be reduced considerably if the computational grids resulting from covering the search domain are non-uniform, i.e.
-when the grid nodes are distributed denser around the global optimal point only.
+when the grid nodes are distributed denser around the global optimal point only (see Fig.~1).
 The construction of such non-uniform coverings may be ensured in the case of an adaptive computation, when in order to select the new trial points, all search information obtained in the course of computation (previous trial points and the values of the minimized function at these points) is used.
@@
 As a result, many optimization algorithms use various dimensionality reduction schemes to certain extent (see, for example, \cite{Strongin1, Strongin2, Sergeyev2}).
+
+\begin{figure}
+\centering
+\includegraphics[height=5.0cm]{Fig1a}
+\includegraphics[height=5.0cm]{Fig1b}
+\caption{Uniform and non-uniform grid}
+\label{fig:Fig1}
+\end{figure}
 
@@
 
+Fig.~2 shows an approximation of the Peano curve mapping a two-dimensional domain with the third mapping density level.
+The accuracy of the Peano curve approximation is determined by its construction density level.
+The order of mapping the two-dimensional domain $D$ is determined by the subdomain connection order resulting from the construction of the Peano curve approximation; the corresponding order of subintervals of the interval $[0, 1]$ is determined by their ordered placement from left to right (i.e.
+the subdomain $\Delta_1$ corresponds to the subinterval $\delta_1$, the subdomain $\Delta_2$ corresponds to the subinterval $\delta_2$, etc).
+
+\begin{figure}
+\centering
+\includegraphics[height=5.0cm]{Fig2}
+\caption{An example of the Peano curve approximation with the third density level}
+\label{fig:Fig2}
+\end{figure}
+
+It should be noted that reducing the multidimensional functions from (1) to the corresponding one-dimensional functions (3) using Peano-type curves preserves the uniform boundedness of the function value differences subject to the bounded argument variations, i.
+e.
+ensures the feasibility of the condition (4).
+Hence the computationally intensive analysis of the multidimensional subdomains to construct a non-uniform coverage of the search domain D can be replaced by a more simplified analysis of reduced one-dimensional data.
+
 \section{Multilevel parallel computational schemes for solving computationally intensive global optimization problems}
+
+A key distinctive feature of the global optimization problems is a limited opportunity to use parallelization methods used widely in the theory and practice or high-performance computing.
+\begin{itemize}
+\item \textit{Data parallelism} (distributing the computational domain among different computing units) is inefficient as in this case only one processor will process the subdomain that contains the sought global minimum while other processors will perform the redundant computations.
+
+\item \textit{Task parallelism} (parallelization of computational schemes) is also inefficient as computational costs of the execution of the optimization algorithm schemes are relatively low (in accordance with the initial supposition, the most time-consuming calculations in the course of global optimization consist in computing the optimized function values).
+
+\item \textit{Parallelization of time-consuming computations} to compute the optimized function values is not general (parallelization methods have to be developed from scratch every time a new optimization problem is solved).
+\end{itemize}
 
@@
 
+It should be noted that the proposed approach can also be applied for solving many other computationally intensive problems -- such as root finding in systems of nonlinear equations, integration of multidimensional integrals, approximation of the data dependencies etc.
+
+In accordance with the approach mentioned above, in this paper a generalized computational scheme of parallel computations to solve the global optimization problems are described as follows.
+To start with, in Subsection 3.1, the general computation scheme of global search algorithms is given.
+Then, in Subsection 3.2, the general scheme of parallel computations for the systems with shared memory is presented.
+In Subsection 3.3, a parallel computational scheme for the systems with distributed memory is described.
+Finally, in Subsection 3.4, a multilevel decomposition scheme for parallel computations for solving the global optimization problems is proposed.
+
 \subsection{General computational scheme of global optimization algorithms}
@@
 The information-statistical theory \cite{Strongin1, Strongin3} formed a basis for many efficient global optimization methods -- see, for example \cite{Strongin2, Sergeyev4, Gergel1, Gergel2, Barkalov, Gergel3, Gergel4, Gergel5, Lera}.
-Moreover, a general computation scheme \cite{Strongin1, Strongin2, Strongin3, Grishagin1, Grishagin2} was proposed for all developed global search algorithms.
+Moreover, a general computation scheme \cite{Strongin1, Strongin2, Strongin3, Grishagin1, Grishagin2} (see also Fig.~3) was proposed for all developed global search algorithms.
 This scheme can be represented as follows.
+
+\begin{figure}
+\centering
+\includegraphics[height=6.0cm]{Fig3}
+\caption{General computation scheme of global optimization algorithms}
+\label{fig:Fig3}
+\end{figure}
 
@@
 
+In order to explain the computational scheme given above, the following can be noted.
 The computed characteristics $R(i), 1 \leq i \leq k + 1$ can be understood as the importance indicators for the intervals with respect to the global minimum point location.
@@
 As it has been noted above, the general approach to provide the parallel computations consists in simultaneous calculating the minimized function values at several different points of the searching domain $D$.
-As the interval characteristics $R(i), 1 \leq i \leq k + 1$ (see Subsection 3.1) are understood as the importance indicators for the intervals with respect to the global minimum point location, additional intervals for several trials to be run in parallel should be selected subject to the following values of the characteristics.
+For the shared memory computing systems, selecting the set of points for executing the parallel trials can be based on the characteristics $R(i), 1 \leq i \leq k + 1$ of the search intervals $(x_{i-1} ,x_i), 1 \leq i \leq k + 1$ introduced in the framework of the general computational scheme (see Subsection 3.1).
+As the characteristics are understood as the importance indicators for the intervals with respect to the global minimum point location, additional intervals for several trials to be run in parallel should be selected subject to the following values of the characteristics.
 
@@
 
-In general, the described general scheme of parallel algorithms of global optimization for systems with shared memory identifies the procedure for competitive parallel computations (see Fig.~1a) where the search intervals $(x_{i-1} ,x_i), 1 \leq i \leq k + 1$ compete with each other for the access to the processors to execute next global search iterations.
+In general, the described general scheme of parallel algorithms of global optimization for systems with shared memory identifies the procedure for competitive parallel computations (see Fig.~4a) where the search intervals $(x_{i-1} ,x_i), 1 \leq i \leq k + 1$ compete with each other for the access to the processors to execute next global search iterations.
 Taking into account the possible number of processors/cores in computational nodes with homogeneous shared memory, the described approach enables the parallelism of about $10^2$ computing units.
@@
 \centering
-\includegraphics[width=13.0cm]{Fig1}
+\includegraphics[width=13.0cm]{Fig4}
 \caption{Generalized computational scheme of parallel algorithms}
-\label{fig:Fig1}
+\label{fig:Fig4}
 \end{figure}
@@
 The conditions of convergence and nonredundancy of the resulting parallel computations for the global optimization algorithms formulated in the framework of general scheme of parallel algorithms of global optimization are given in \cite{Strongin1, Strongin2, Strongin4, Grishagin2}.
+Thus, in particular, when the convergence conditions are met, the parallel computing for the MGAGS algorithm is non-redundant involving up to $2^N$ processors (cores) where $N$ is the dimension of the global optimization problem to be solved.
 
@@
 
-Using such a parallelization approach can result in the following general scheme of parallel algorithms of global optimization for the systems with distributed and shared memory -- see also \cite{Gergel1} and Fig.~1b.
+Using such a parallelization approach can result in the following general scheme of parallel algorithms of global optimization for the systems with distributed and shared memory -- see also \cite{Gergel1} and Fig.~4b.
 
@@
 \end{enumerate}
+
+It should be noted that this parallel computing scheme does not have any central control node.
+The number of computational nodes may vary during the global optimization, and removing any node will not lead to the loss in finding the global minimum of the optimization problem (1).
+In addition, the data communication may be asynchronous (the computational nodes should not wait until the optimization data are transmitted).
 
@@
 
-To enable the parallel computations, numerous nested optimization subproblems can be generated at once on each decomposition level for solving these ones in parallel \cite{Sergeyev7} (see Fig.~1c).
+To enable the parallel computations, numerous nested optimization subproblems can be generated at once on each decomposition level for solving these ones in parallel \cite{Sergeyev7} (see Fig.~4c).
 The resulting number of parallel subproblems can be controlled using a predefined parallelization vector
@@
 
-To obtain more reliable estimates of the parallelization efficiency, 100 test problems of global optimization were solved. The maximum allowed number of parallel iterations was $Kmax = 10\,000$.
+To obtain more reliable estimates of the parallelization efficiency, 100 test problems of global optimization were solved.
+The global minima $y^\ast$ of the test problems were considered to be found if a trial point $y^k$ is generated in the $\delta$-vicinity of the global minimum, i.e.
+$\| y^k - y^ \ast \| \leq \delta$.
+The vicinity size was set equal to $\delta = 0.01 \|b - a \|$ , where $a$ and $b$ were the boundaries of the search domain $D$.
+The maximum allowed number of parallel iterations was $Kmax = 10\,000$.
 
@@
 The proposed schemes ensure the computational efficiency of the supercomputing systems with shared and distributed memory with thousands of processors to address the global optimization challenges.
+The executed computational experiments confirm the potential of the proposed approach.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{acknowledgments}
+This research was supported by the Russian Science Foundation, project No 16-11-10150 "Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance".
+\end{acknowledgments}
 

### 2018-01-18 2017/KAZAN_LJM/paper_lobach.tex
some corrections

@@
 \titlerunning{Generalized Parallel Schemes for Global Optimization}
-\authorrunning{Strongin, Gergel, Barkalov, Sysoyev}
+\authorrunning{Strongin et al.}
 
@@
 \email[E-mail: ]{president@unn.ru}
-\affiliation{Lobachevsky State University of Nizhni Novgorod}
+\affiliation{Lobachevsky State University of Nizhni Novgorod, pr. Gagarina 23, Nizhni Novgorod, Russia}
 
@@
 \email[E-mail: ]{gergel@unn.ru}
-\affiliation{Lobachevsky State University of Nizhni Novgorod}
+\affiliation{Lobachevsky State University of Nizhni Novgorod, pr. Gagarina 23, Nizhni Novgorod, Russia}
 
@@
 \email[E-mail: ]{barkalov@vmk.unn.ru}
-\affiliation{Lobachevsky State University of Nizhni Novgorod}
+\affiliation{Lobachevsky State University of Nizhni Novgorod, pr. Gagarina 23, Nizhni Novgorod, Russia}
 
@@
 \email[E-mail: ]{sysoyev@vmk.unn.ru}
-\affiliation{Lobachevsky State University of Nizhni Novgorod}
+\affiliation{Lobachevsky State University of Nizhni Novgorod, pr. Gagarina 23, Nizhni Novgorod, Russia}
 
@@
 
-The state-of-the-art in the field of global optimization is outlined in a number of key works – see, for instance, \cite{Floudas, Locatelli, Strongin1, Pardalos, Sergeyev1, Paulavicius}.
+The state-of-the-art in the field of global optimization is outlined in a number of key works -- see, for instance, \cite{Floudas, Locatelli, Strongin1, Pardalos, Sergeyev1, Paulavicius}.
 The development of the theory and methods of use of the high-performance computing for solving the computationally intensive problems of global optimization is one of the trends developing rapidly \cite{Strongin1, Ciegis, Luque, Strongin2}.
@@
 In the framework of the information-statistical theory, the dimensionality reduction is provided by using the Peano space-filling curves $y(x)$, which map the interval $[0, 1]$ onto a $N$-dimensional domain $D$ \cite{Strongin1, Sergeyev2, Strongin3}.
-As a result of such reduction, the initial multidimensional global optimization problem (1) is reduced to a one-dimensional one:
+As a result of such reduction, the initial multidimensional global optimization problem (1) is reduced to a one-dimensional one
 \begin{equation}
-\varphi(y(x^\ast)) = \min\{\varphi(y(x)): x \in [0,1]\}.
+\varphi(y(x^\ast)) = \min\{\varphi(y(x)): x \in [0,1]\},
 \end{equation}
@@
 
-It should be noted that the proposed approach can also be applied for solving many other computationally intensive problems -- such as root finding in systems of nonlinear equations, integration of multidimensional integrals, approximation of the data dependencies etc.
+It should be noted that the proposed approach can also be applied for solving many other computationally intensive problems such as root finding in systems of nonlinear equations, integration of multidimensional integrals, approximation of the data dependencies etc.
 
@@
 \begin{equation}
-(x_t - x_{t-1}) \leq \varepsilon,
+(x_t - x_{t-1}) \leq \varepsilon
 \end{equation}
-for the interval $t, 1 \leq i \leq k + 1$ with the maximum characteristic from (6) and $\varepsilon > 0$ is the required problem solving accuracy.
+for the interval $t$ with the maximum characteristic from (6); $\varepsilon > 0$ is the required problem solving accuracy.
 If the stopping condition is not satisfied, the iteration number $k$ is incremented by unity, and a new global search iteration is executed.
@@
 In order to explain the computational scheme given above, the following can be noted.
-The computed characteristics $R(i), 1 \leq i \leq k + 1$ can be understood as the importance indicators for the intervals with respect to the global minimum point location.
+The computed characteristics $R(i), 1 \leq i \leq k + 1$, can be understood as the importance indicators for the intervals with respect to the global minimum point location.
 This provides the explanation for the interval selection scheme for executing the next trial: the point of every next global search iteration is selected in the interval with the maximum characteristic (i.e.
@@
 \end{equation}
-where $m$ is the H\"older constant estimate from (4), $z_i = \varphi(y(x_i)), 1 \leq i \leq k + 1$ are the computed values of the minimized function $\varphi(y(x))$ at the points of the executed iterations, and $\rho_i = \left(x_i - x_{i-1} \right)^{1/N}, 1 \leq i \leq k + 1$.
+where $m$ is the H\"older constant estimate from (4), $z_i = \varphi(y(x_i)), 1 \leq i \leq k + 1$, are the computed values of the minimized function $\varphi(y(x))$ at the points of the executed iterations, and $\rho_i = \left(x_i - x_{i-1} \right)^{1/N}, 1 \leq i \leq k + 1$.
 
@@
 \end{equation}
-where $r, r > 1$ is a \textit{parameter} of the algorithm.
+where $r > 1$ is a \textit{parameter} of the algorithm.
 
@@
 As it has been noted above, the general approach to provide the parallel computations consists in simultaneous calculating the minimized function values at several different points of the searching domain $D$.
-For the shared memory computing systems, selecting the set of points for executing the parallel trials can be based on the characteristics $R(i), 1 \leq i \leq k + 1$ of the search intervals $(x_{i-1} ,x_i), 1 \leq i \leq k + 1$ introduced in the framework of the general computational scheme (see Subsection 3.1).
+For the shared memory computing systems, selecting the set of points for executing the parallel trials can be based on the characteristics $R(i), 1 \leq i \leq k + 1$, of the search intervals $(x_{i-1} ,x_i), 1 \leq i \leq k + 1$, introduced in the framework of the general computational scheme (see Subsection 3.1).
 As the characteristics are understood as the importance indicators for the intervals with respect to the global minimum point location, additional intervals for several trials to be run in parallel should be selected subject to the following values of the characteristics.
 
-As a result, a general scheme of the parallel algorithms of global optimization for systems with shared memory can be formulated on the basis of the scheme given in Subsection 3.1 where the following set of rules should be modified (see also \cite{Strongin1, Strongin2, Strongin4, Grishagin2}):
+As a result, a general scheme of the parallel algorithms of global optimization for systems with shared memory can be formulated on the basis of the scheme given in Subsection 3.1 where the following set of rules should be modified (see also \cite{Strongin1, Strongin2, Strongin4, Grishagin2}).
 
@@
 \end{equation}
-and select $p$ intervals with the numbers $t_j, 1 \leq j \leq p$ having the maximum values of characteristics ($p$ is the number of processors (cores) used for the parallel computing).
+and select $p$ intervals with the numbers $t_j, 1 \leq j \leq p$, having the maximum values of characteristics ($p$ is the number of processors (cores) used for the parallel computing).
 
 \textit{Rule} $4'$.
-Execute new trials (compute the values of the minimized function $\varphi(y(x))$) at the points $x^{k+j}, 1 \leq j \leq p$ located within the intervals with the maximum characteristics from (10).
+Execute new trials (compute the values of the minimized function $\varphi(y(x))$) at the points $x^{k+j}, 1 \leq j \leq p$, located within the intervals with the maximum characteristics from (10).
 
@@
 
-In general, the described general scheme of parallel algorithms of global optimization for systems with shared memory identifies the procedure for competitive parallel computations (see Fig.~4a) where the search intervals $(x_{i-1} ,x_i), 1 \leq i \leq k + 1$ compete with each other for the access to the processors to execute next global search iterations.
+In general, the described general scheme of parallel algorithms of global optimization for systems with shared memory identifies the procedure for competitive parallel computations (see Fig.~4a) where the search intervals $(x_{i-1} ,x_i), 1 \leq i \leq k + 1$, compete with each other for the access to the processors to execute next global search iterations.
 Taking into account the possible number of processors/cores in computational nodes with homogeneous shared memory, the described approach enables the parallelism of about $10^2$ computing units.
@@
 \begin{equation}
-Y_s(x) = \{ y^1(x), \dots , y^s(x) \},
+Y_s(x) = \{ y^1(x), \dots , y^s(x) \}
 \end{equation}
@@
 $ y^\ast = y^l(x_l^\ast), 1 \leq l \leq s$.
-It is important to note that a family of the one-dimensional problems $\phi(y^l(x)), 1 \leq l \leq s$ resulting from the dimensionality reduction is an information-linked one: the function values computed for some function $\phi(yl(x))$ can be used by all other problems of the family (14).
+It is important to note that a family of the one-dimensional problems $\phi(y^l(x)), 1 \leq l \leq s$, resulting from the dimensionality reduction is an information-linked one: the function values computed for some function $\phi(yl(x))$ can be used by all other problems of the family (14).
 
@@
 \begin{itemize}
-\item Before starting a new trial for any problem $\phi(y^l(x)), 1 \leq l \leq s$ at any point $x' \in [0,1]$, it is necessary to calculate the image $y' \in D$ of the point $x' \in [0,1]$ according to the mapping of $y^l(x)$ and compute the preimages of $x_i', 1 \leq i \leq s$ for each problem of the family $\overrightarrow{\Phi}(y)$.
-The resulting preimages $x_i', 1 \leq i \leq s$ should be sent to all involved computational nodes to eliminate repeated selection of new trial points in the intervals containing the preimages.
+\item Before starting a new trial for any problem $\phi(y^l(x)), 1 \leq l \leq s$, at any point $x' \in [0,1]$, it is necessary to calculate the image $y' \in D$ of the point $x' \in [0,1]$ according to the mapping of $y^l(x)$ and compute the preimages of $x_i', 1 \leq i \leq s$, for each problem of the family $\overrightarrow{\Phi}(y)$.
+The resulting preimages $x_i', 1 \leq i \leq s$, should be sent to all involved computational nodes to eliminate repeated selection of new trial points in the intervals containing the preimages.
 
-\item Upon the completing of any trial for any problem $\phi(y^l(x)), 1 \leq l \leq s$ at any point $x' \in [0,1]$, calculate all preimages $x_i', 1 \leq i \leq s$ again and distribute these preimages and the trial result $z' = \phi(y^l(x'))$ among all involved computational nodes to include the obtained data in the search information (5).
+\item Upon the completing of any trial for any problem $\phi(y^l(x)), 1 \leq l \leq s$, at any point $x' \in [0,1]$, calculate all preimages $x_i', 1 \leq i \leq s$, again and distribute these preimages and the trial result $z' = \phi(y^l(x'))$ among all involved computational nodes to include the obtained data in the search information (5).
 
@@
 \[
-y = (y_1, y_2, ..., y_N) = (u_1, u_2, ..., u_M)
+y = (y_1, y_2, ..., y_N) = (u_1, u_2, ..., u_M),
 \]
@@
 u_1 = (y_1, y_2, ..., y_{N_1}), \dots, 
-u_M = (y_{N-N_M+1}, y_{N-N_M+2}, ..., y_N).
+u_M = (y_{N-N_M+1}, y_{N-N_M+2}, ..., y_N),
 \]
@@
 
-The computational experiments were performed using the Lobachevsky supercomputer at the State University of Nizhny Novgorod (operating system -- CentOS 6.4, supercomputer management system - SLURM).
-Each computational node had two 8-core Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 GB RAM, and 3 NVIDIA Kepler K20Х graphic processors ($2\,688$ cores, 6 GB GDDR5 RAM).
+The computational experiments were performed using the Lobachevsky supercomputer at the State University of Nizhni Novgorod (CentOS 6.4 operating system, SLURM supercomputer management system).
+Each computational node had two 8-core Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 GB RAM, and 3 NVIDIA Kepler K20X graphic processors ($2\,688$ cores, 6 GB GDDR5 RAM).
 Intel C++ 14.0.2 was used to generate the executable program code.
@@
 The vicinity size was set equal to $\delta = 0.01 \|b - a \|$ , where $a$ and $b$ were the boundaries of the search domain $D$.
-The maximum allowed number of parallel iterations was $Kmax = 10\,000$.
+The maximum allowed number of parallel iterations was $K_{max} = 10\,000$.
 
-The computational experiments involved the use of $p$ nodes of the Lobachevsky supercomputer, each node used three NVIDIA Kepler K20Х graphic accelerators.
+The computational experiments involved the use of $p$ nodes of the Lobachevsky supercomputer, each node used three NVIDIA Kepler K20X graphic accelerators.
 Thus, for $p = 32$, the number of GPU accelerators were 96, each having $2\,688$ CUDA cores (i.e.
@@
 \begin{acknowledgments}
-This research was supported by the Russian Science Foundation, project No 16-11-10150 "Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance".
+This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''.
 \end{acknowledgments}
@@
 \bibitem{Strongin4}
-R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global multidimensional optimization on parallel computer}, Parallel Computing, \textbf{18} (11), 1259--1273, (1992).
+R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global multidimensional optimization on parallel computer}, Parallel Computing \textbf{18} (11), 1259--1273 (1992).
 
 \bibitem{Sergeyev4}
-Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel asynchronous global search and the nested optimization scheme}, J. Comput. Anal. Appl., \textbf{3} (2), 123--145, (2001).
+Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel asynchronous global search and the nested optimization scheme}, J. Comput. Anal. Appl., \textbf{3} (2), 123--145 (2001).
 
 \bibitem{Gergel1}
-V.~P.~Gergel and S.~V.~Sidorov, \textit{A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems}, V.
-Malyshkin (Ed.): PaCT 2015, LNCS 9251, 505--515.
-Springer-Verlag Berlin Heidelberg ,(2015).
+V.~P.~Gergel and S.~V.~Sidorov, \textit{A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems}, Lecture Notes in Computer Science \textbf{9251}, 505--515 (2015).
 
 \bibitem{Gergel2}
-V.~Gergel, \textit{An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems}, 2nd International Conference on Mathematics and Computers in Sciences and in Industry, 13--18, (2015).
+V.~Gergel, \textit{An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems}, in Proceedings of the Second International Conference on Mathematics and Computers in Sciences and in Industry (MCSI), Sliema, 13--18 (2015).
 
 \bibitem{Barkalov}
-K.~Barkalov, V.~Gergel and I.~Lebedev, \textit{Solving global optimization problems on GPU cluster}, In: Simos, T.E. (ed.) ICNAAM 2015. AIP Conference Proceedings, 1738, art. no. 400006, (2016).
+K.~Barkalov, V.~Gergel and I.~Lebedev, \textit{Solving global optimization problems on GPU cluster}, AIP Conference Proceedings \textbf{1738}, 400006 (2016).
 
 \bibitem{Gergel3}
-V.~Gergel and E.~Kozinov, \textit{Efficient methods of multicriterial optimization based on the intensive use of search information}, Springer Proceedings in Mathematics and Statistics, 197, 27--45, (2017).
+V.~Gergel and E.~Kozinov, \textit{Efficient methods of multicriterial optimization based on the intensive use of search information}, Springer Proceedings in Mathematics and Statistics \textbf{197}, 27--45 (2017).
 
 \bibitem{Gergel4}
-V.~Gergel and E.~Kozinov, \textit{Parallel computing for time-consuming multicriterial optimization problems}, Lecture Notes in Computer Science, 10421, 446--458, (2017).
+V.~Gergel and E.~Kozinov, \textit{Parallel computing for time-consuming multicriterial optimization problems}, Lecture Notes in Computer Science \textbf{10421}, 446--458 (2017).
 
 \bibitem{Gergel5}
-V.~Gergel, V.~Grishagin and A.~Gergel, \textit{Adaptive nested optimization scheme for multidimensional global search}, Journal of Global Optimization, \textbf{66} (1), 35--51, (2016).
+V.~Gergel, V.~Grishagin and A.~Gergel, \textit{Adaptive nested optimization scheme for multidimensional global search}, J. Glob. Optim.  \textbf{66} (1), 35--51 (2016).
 
 \bibitem{Lera}
-D.~Lera, Y.~D.~Sergeyev, \textit{Lipschitz and Holder global optimization using space-filling curves}, Appl. Numer. Math, \textbf{60} (1-2), 115--129, (2010).
+D.~Lera, Y.~D.~Sergeyev, \textit{Lipschitz and Holder global optimization using space-filling curves},  Appl. Numer. Math. \textbf{60} (1-2), 115--129 (2010).
 
 \bibitem{Grishagin1}
-V~A.~Grishagin, \textit{On convergence conditions for a class of global search algorithms}, Proceedings of the 3-rd All-Union seminar "Numerical methods of nonlinear programming", Kharkov, 82--84, (1979) [In Russian].
+V~A.~Grishagin.
+\textit{On convergence conditions for a class of global search algorithms}, in Proceedings of the 3-rd All-Union seminar ``Numerical methods of nonlinear programming'', Kharkov, 82--84 (1979) [In Russian].
 
 \bibitem{Grishagin2}
-V.~A.~Grishagin, Y.~D.~Sergeyev and R.~G.~Strongin, \textit{Parallel characteristic algorithms for solving problems of global optimization}, Journal of Global Optimization, \textbf{10} (2), 185--206, (1997).
+V.~A.~Grishagin, Y.~D.~Sergeyev and R.~G.~Strongin, \textit{Parallel characteristic algorithms for solving problems of global optimization}, J. Glob. Optim. \textbf{10} (2), 185--206 (1997).
 
 \bibitem{Strongin5}
-R.~G.~Strongin, \textit{Algorithms for Multi-extremal Mathematical Programming Problems Employing the Set of Joint Space-filling Curves}, Journal of Global Optimization, \textbf{2} (4), 357--378, (1992).
+R.~G.~Strongin, \textit{Algorithms for Multi-extremal Mathematical Programming Problems Employing the Set of Joint Space-filling Curves}, J. Glob. Optim. \textbf{2} (4), 357--378 (1992).
 
 \bibitem{Sysoyev}
-A.~Sysoyev, K.~Barkalov, V.~Sovrasov, I.~Lebedev and V.~Gergel, \textit{Globalizer –- A parallel software system for solving global optimization problems}, Lecture Notes in Computer Science 10421, 492--499, (2017).
+A.~Sysoyev, K.~Barkalov, V.~Sovrasov, I.~Lebedev and V.~Gergel, \textit{Globalizer -- A parallel software system for solving global optimization problems}, Lecture Notes in Computer Science \textbf{10421}, 492--499 (2017).
 
 \bibitem{Sergeyev7}
-Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel Asynchronous Global Search and the Nested Optimization Scheme}, J. Comput. Anal. Appl., \textbf{3} (2), 123--145, (2001).
+Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel Asynchronous Global Search and the Nested Optimization Scheme}, J. Comput. Anal. Appl. \textbf{3} (2), 123--145 (2001).
 
@@
 Math.
-Software, \textbf{29}, 469--480, (2003).
+Softw.
+\textbf{29}, 469--480 (2003).
 
 \bibitem{Modorskii}
-V.~Y.~Modorskii, D.~F.~Gaynutdinova, V.~P.~Gergel and K.~A.~Barkalov, \textit{Optimization in design of scientific products for purposes of cavitation problems}, AIP Conference Proceedings, 1738, art.
-no.
-400013, (2016).
+V.~Y.~Modorskii, D.~F.~Gaynutdinova, V.~P.~Gergel and K.~A.~Barkalov, \textit{Optimization in design of scientific products for purposes of cavitation problems}, AIP Conference Proceedings \textbf{1738}, 400013 (2016).
 
 \bibitem{Gergel6}
-V.~P.~Gergel, M.~I.~Kuzmin, N.~A.~Solovyov and V.~A.~Grishagin, \textit{Recognition of surface defects of cold-rolling sheets based on method of localities}, International Review of Automatic Control, \textbf{8} (1), 51--55, (2015).
+V.~P.~Gergel, M.~I.~Kuzmin, N.~A.~Solovyov and V.~A.~Grishagin, \textit{Recognition of surface defects of cold-rolling sheets based on method of localities}, International Review of Automatic Control \textbf{8} (1), 51--55 (2015).
 

### 2018-01-24 2018/ITNT/ITNT2018.tex
Final version

@@
 
-In the algorithms considered in the present paper the objective function and constraints are assumed to be Lipschitzian that is typical for many other approaches to the development of the parallel optimization algorithms (see, for example, \cite{Jones,Zilinskas,Evtushenko}).
+In the algorithms considered in the present paper the objective function and constraints are assumed to be Lipschitzian that is typical for many other approaches to the development of the parallel optimization algorithms (see, for example, \cite{Jones,Zilinskas,Evtushenko2009,Evtushenko2014}).
 This assumption is natural for many applied problems since the relative variations of the functions generally cannot exceed a certain threshold determined by the bounded energy of changes occurring in the system under study.
@@
 It is assumed that it is possible to select the ``difficult'' and ``easy'' parts 
-of the functions connected with each other. Then, the solving of the 
-``difficult'' part of the problem (for which the execution of time-consuming algorithms is required to conduct the trials) can be performed on CPU using an efficient index global optimization algorithm.
-The ``easy'' part
-of the problem (the execution of which doesn't require the complex algorithms 
+of the functions connected with each other. An example of a problem of this type has been considered in \cite{Strongin2017}.
+Then, the solving of the ``difficult'' part of the problem (for which the execution of time-consuming algorithms is required to conduct the trials) can be performed on CPU using an efficient index global optimization algorithm.
+The ``easy'' part of the problem (the execution of which doesn't require the complex algorithms
 and can be transferred to an accelerator easily) can be solved on GPU using 
@@
 with search domain
-\begin{equation}\label{D}
+\[
 D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.
 %D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.
-\end{equation}
+\]
 
@@
 the left-hand sides $g_i(y), \; 1\leq i \leq m$, of the constraints
-satisfy the Lipschitz conditions with constants $L_i, \; 1 \leq i \leq
+satisfy the Lipschitz conditions with constants $L_i$, $1 \leq i \leq
 m+1$, respectively, and may be multiextremal. It is assumed that the
@@
 The use of Peano curve $y(x)$ 
-\begin{equation}
+\[
 \left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq 
 N\right\}=\left\{y(x):0\leq x \leq 1 \right\}
-\end{equation}
+\]
 unambiguously mapping the interval of real axis $[0,1]$ onto a $N$-dimensional cube is the first of the dimensionality reduction methods considered.
@@
 Nested optimization scheme is based on relation (see \cite{Strongin2013})
-\begin{equation}\label{nested}
+\[
 \min_{y \in D}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq 
@@
 }{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
-\end{equation}
+\]
 which allows replacing the solving of multidimensional problem 
@@
 Here we consider vector $y$ as a vector of block variables
-\begin{equation}
+\[
 y=(y_1,y_2,...,y_N)=(u_1,u_2,...,u_M),
-\end{equation}
+\]
 where the $i$-th block variable $u_i$ is a vector of $N_i$ components of 
@@
 nesting level has been used
-\begin{equation}
+\[
 \min_{y \in D}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq 
@@
 g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}.
-\end{equation}
+\]
 
@@
 
-\bibitem{Evtushenko}
+\bibitem{Evtushenko2009}
 Evtushenko, Yu.G. Parallel global optimization of functions of several 
@@
 (2). --- P. 246--260.
+
+\bibitem{Evtushenko2014}
+Evtushenko, Yu.G.
+Method of non-uniform coverages to solve the multicriteria optimization problems with guaranteed accuracy / Yu.G.
+Evtushenko, M.
+A.
+Posypkin // Automation and Remote Control.
+--- 2014.
+--- Vol.
+75(6).
+--- P.
+1025--1040.
+
+\bibitem{Strongin2017}
+Barkalov, K.A. Test problems for parallel algorithms of constrained global optimization / K.A. Barkalov, R.G. Strongin // Lecture Notes in Computer Science. --- 2017. --- Vol. 10556. --- P. 18--33.
 

### 2018-01-24 2018/PAVT_Toropov/PAVT2018_MAM.tex
Final version

@@
 
-The Moving Least Squares Method (MLSM) was proposed in \cite{LancasterSalkauskas1981} for smoothing and interpolation of scattered data and later used in the mesh-free form of the FEM \cite{Liszka1984}.
-As suggested in \cite{ChoiYounYang2001}, it can be used as a technique for metamodelling and used in MDO frameworks.
+The Moving Least Squares Method (MLSM) was proposed in \cite{LancasterSalkauskas1981} for smoothing and interpolation of scattered data and later used in the mesh-free form of the finite element method (FEM) \cite{Liszka1984}.
+As suggested in \cite{ChoiYounYang2001}, it can be used as a technique for metamodelling and used in multidisciplinary optimization (MDO) frameworks.
 The MLSM is a weighted least squares method where the weights depend on the Euclidean distance from a sample point to where the surrogate model is to be evaluated.
@@
 Toropov, V., Filatov, A., Polynkin, A.:
-\newblock Multiparameter structural optimization using fem and multipoint explicit approximations.
+\newblock Multiparameter structural optimization using FEM and multipoint explicit approximations.
 \newblock Structural Optimization \textbf{6}, 7--14 (1993)
@@
 \newblock Large scale optimization of transonic axial compressor rotor blades.
-\newblock In:  49th AIAA/ASME/ASCE/AHS/ASC Structures, Structural Dynamics, and Materials Conference, AIAA, art. no. 2008-2056
- (2008)
+\newblock In:  49th AIAA/ASME/ASCE/AHS/ASC Structures, Structural Dynamics, and Materials Conference, AIAA, art. no. 2008-2056 (2008)
 
@@
 
-%\bibitem{Strongin2000}
-%Strongin, R.G., Sergeyev, Y.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000); DOI: 10.1007/978-1-4615-4677-1
-%
-%\bibitem{Barkalov2002}
-%Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)
-%
-%\bibitem{Gergel2005}
-%Gergel, V.P., Strongin, R.G.: Parallel computing for globally optimal decision making on cluster systems. Future Gener. Comput. Syst. 21(5), 673--678 (2005); DOI: 10.1016/j.future.2004.05.007
-%
-%\bibitem{Strongin2013}
-%Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013);  DOI: 10.1007/978-1-4614-8042-6
-%
-%\bibitem{Barkalov2010}
-%Barkalov, K., Ryabov, V., Sidorov, S.: Parallel Scalable Algorithms with Mixed Local-Global Strategy for Global Optimization Problems.
-In: Hsu, Ch., Malyshkin, V.
-(Eds.) MTPP 2010, LNCS, vol.
-6083, pp.
-232--240.
-Springer, Heidelberg (2010); DOI: 10.1007/978-3-642-14822-4$\_$26
-%
-%\bibitem{Barkalov2015}
-%Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi Coprocessor for Solving Global Optimization Problems.
-In: Malyshkin, V.
-(Ed.) PaCT 2015, LNCS, vol.
-9251, pp.
-307--318.
-Springer, Heidelberg (2015); DOI: 10.1007/978-3-319-21909-7$\_$31
-%
-%\bibitem{Barkalov2016}
-%Barkalov, K., Gergel, V.: Parallel Global Optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016); DOI: 10.1007/s10898-016-0411-y
-%
-%\bibitem{Gergel2016}
-%Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster.
-In: Simos T.E.
-(Ed.) ICNAAM 2015, AIP Conference Proceedings, vol.
-1738, art.
-no.
-400006 (2016); DOI: 10.1063/1.4952194
-%
-%\bibitem{Gergel2015}
-%Gergel, V., Sidorov, S.: A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems.
-In: Malyshkin, V.
-(Ed.) PaCT 2015, LNCS, vol.
-9251, pp.
-505--515.
-Springer, Heidelberg (2015); DOI: 10.1007/978-3-319-21909-7$\_$49
-%
-
-%\end{thebibliography}
-
 \end{document}

### 2018-01-25 2018/PAVT_Toropov/PAVT2018_MAM.tex
final version

@@
 nonlinear functions that can be led to a linear form by a simple transformation) for individual sub-structures, and assembly of them for the whole structure.
-This was enhanced by the use of gradient-assisted metamodels \cite{ToropovFilatov1993}, use of simplified numerical models that is also termed a multi-fidelity approach,\cite{ToropovMarkine1996} and the use of analytical models derived by Genetic Programming \cite{ToropovAlvarez1998}.
+This was enhanced by the use of gradient-assisted metamodels \cite{ToropovFilatov1993}, use of simplified numerical models that is also termed a multi-fidelity approach \cite{ToropovMarkine1996}, and the use of analytical models derived by Genetic Programming \cite{ToropovAlvarez1998}.
 One of the recent developments \cite{PolynkinToropov2012} involved the use of approximation assemblies, i.e.

### 2018-03-16 2017/NACO/NACO-16-0035_3.tex

--- 2017-06-27 2017/NACO/paper.tex
@@
 %local folder where your tex file resides. http://aimsciences.org/journals/tex-sample/AIMS.cls.
-\documentclass{aims}
+\documentclass{naco}
 \usepackage{amsmath}
@@
 
-\usepackage[utf8]{inputenc}
+%\usepackage[utf8]{inputenc}
 
@@
     \topmargin 30pt
-     \setcounter{page}{1}
+     \setcounter{page}{47}
 
 % The next 5 line will be entered by an editorial staff.
-\def\currentvolume{X}
- \def\currentissue{X}
-  \def\currentyear{200X}
-   \def\currentmonth{XX}
-    \def\ppages{X--XX}
-     \def\DOI{10.3934/xx.xx.xx.xx}
-
+\def\currentvolume{8}
+ \def\currentissue{1}
+  \def\currentyear{2018}
+   \def\currentmonth{March}
+    \def\ppages{47--62}
+     \def\DOI{10.3934/naco.2018003}
  % Please minimize the usage of "newtheorem", "newcommand", and use
@@
 
-\centerline{\scshape Victor Gergel, Konstantin  Barkalov$^*$ and Alexander Sysoyev}
+\centerline{\scshape Victor Gergel, Konstantin  Barkalov$^*$  }
+\centerline{\scshape and Alexander Sysoyev}
 \medskip
@@
 % "Communicated by the associate editor name" is not needed for special issue.
- \centerline{(Communicated by the associate editor name)}
+ \centerline{(Communicated by Yaroslav Sergeyev)}
 
@@
 \par
-The list provided above is certainly not exhaustive – additional information on the
+The list provided above is certainly not exhaustive --- additional information on the
 software systems for solving a wider spectrum of optimization problems can be found, for example, in
@@
   solved using the multicriterial optimization block, which, in turn, uses the nonlinear programming block, and so on.
-  \item Block 5 is a subsystem for accumulating and processing the search information; it is one of the main subsystems – the size of search information may appear to be quite large and the efficiency of the global optimization methods depends on how completely the available search data are utilized.
+  \item Block 5 is a subsystem for accumulating and processing the search information; it is one of the main subsystems --- the size of search information may appear to be quite large and the efficiency of the global optimization methods depends on how completely the available search data are utilized.
   \item Block 6 contains the dimensionality reduction procedures based on the Peano space-filling curves; the reduced (one-dimensional) optimization problems are solved by the optimization blocks.
@@
 The numerical experiments were carried out using the Lobachevsky supercomputer at
-the State University of Nizhni Novgorod (the operation system – CentOS 6.4, the
-supercomputer management system – SLURM). Each supercomputer node had 2 Intel Sandy Bridge
+the State University of Nizhni Novgorod (the operation system --- CentOS 6.4, the
+supercomputer management system --- SLURM). Each supercomputer node had 2 Intel Sandy Bridge
 E5-2660 2.2 GHz processors with 64 GB RAM. Each processor had 8 cores (i. e. total
@@
 
-\begin{thebibliography}{10}
+\begin{thebibliography}{99}
 
@@
 \newblock K. Barkalov and V. Gergel,
-\newblock Multilevel scheme of dimensionality reduction for parallel global search algorithms,
-\newblock in \emph{Proceedings of the 1st International Conference on
-  Engineering and Applied Sciences Optimization}, (2014), 2111--2124.
-
-\bibitem{barkalovGergel2015}[10.1007/s10898-016-0411-y]
+\newblock \emph{\emph{Multilevel scheme of dimensionality reduction for parallel global search algorithms}},
+\newblock in \emph{Proceedings of the 1st International Conference on Engineering and Applied Sciences Optimization}, (2014), 2111--2124.
+
+\bibitem{barkalovGergel2015} (MR3539505) [10.1007/s10898-016-0411-y]
 \newblock K. Barkalov and V. Gergel,
-\newblock Parallel global optimization on GPU,
-\newblock \emph{J. Glob. Optim.}, \textbf{66(1)} (2016), 3--20.
-
-\bibitem{barkalovGergelLebedev2015}[10.1007/978-3-319-21909-7_31]
+\newblock \emph{\emph{Parallel global optimization on GPU}},
+\newblock \emph{J. Glob. Optim.}, \textbf{66} (2016), 3--20.
+
+\bibitem{barkalovGergelLebedev2015} [10.1007/978-3-319-21909-7_31]
 \newblock K. Barkalov, V. Gergel and I. Lebedev,
-\newblock Use of Xeon Phi coprocessor for solving global optimization problems,
+\newblock \emph{\emph{Use of Xeon Phi coprocessor for solving global optimization problems}},
 \newblock \emph{LNCS}, \textbf{9251} (2015), 307--318.
@@
 \bibitem{barkalovGergelLebedevSysoev2015}
-\newblock K. Barkalov, V. Gergel, I. Lebedev, A. Sysoev, %http://ceur-ws.org/Vol-1482/
-\newblock Solving the global optimization problems on heterogeneous cluster systems,
+\newblock K. Barkalov, V. Gergel, I. Lebedev and A. Sysoev,
+\newblock \emph{\emph{Solving the global optimization problems on heterogeneous cluster systems}},
 \newblock in \emph{CEUR Workshop Proceedings}, \textbf{1482} (2015), 411--419.
-%\newblock (In Russian).
-
-\bibitem{Barkalov2013}[10.1007/978-3-642-39958-9_14]
-\newblock K. Barkalov, A. Polovinkin, I. Meyerov, S. Sidorov, N. Zolotykh,
-\newblock SVM regression parameters optimization using parallel global search algorithm,
+
+\bibitem{Barkalov2013} [10.1007/978-3-642-39958-9_14]
+\newblock K. Barkalov, A. Polovinkin, I. Meyerov, S. Sidorov and N. Zolotykh,
+\newblock \emph{\emph{SVM regression parameters optimization using parallel global search algorithm}},
 \newblock \emph{LNCS}, \textbf{7979} (2013), 154--166.
 
-\bibitem{bussieckMeeraus2004}[10.1007/978-1-4613-0215-5_8]
+\bibitem{bussieckMeeraus2004} (MR3618583) [10.1007/978-1-4613-0215-5_8]
 \newblock M. R. Bussieck and A. Meeraus,
-\newblock General algebraic modeling system (GAMS), in \emph{Modeling Languages in Mathematical Optimization} (ed. J. Kallrath),
-\newblock Springer (2004), 137--157.
-
-\bibitem{censorZeniosParGO}
+\newblock \emph{\emph{General algebraic modeling system (GAMS)}},
+\newblock in \emph{Modeling Languages in Mathematical Optimization} (ed. J. Kallrath), Springer, (2004), 137--157.
+
+\bibitem{censorZeniosParGO} (MR1486040)
 \newblock Y. Censor and S. A. Zenios,
-\newblock \emph{Parallel optimization: theory, algorithms, and applications},
-\newblock Oxford University Press (1998).
-
-\bibitem{ciegisHentyParGO}[10.1007/978-0-387-09707-7]
+\newblock \emph{Parallel Optimization: Theory, Algorithms, and Applications},
+\newblock Oxford University Press, 1998.
+
+\bibitem{ciegisHentyParGO} (MR2499546) [10.1007/978-0-387-09707-7]
 \newblock R. \v Ciegis, D. Henty, B. K\aa gstr\"om and J. \v Zilinskas,
-\newblock \emph{Parallel scientific computing and optimization: advances and applications},
-\newblock Springer (2009).
+\newblock \emph{Parallel Scientific Computing and Optimization: Advances and Applications},
+\newblock Springer, 2009.
 
@@
 \newblock I. N. Egorov, G. V. Kretinin, I. A. Leshchenko and S. V. Kuptzov,
-\newblock IOSO optimization toolkit --- novel software to create better design,
-\newblock in \emph{9th AIAA/ISSMO Symposium on Multidisciplinary Analysis and Optimization}, 2002.
-\newblock Available from \url{http://www.iosotech.com/text/2002\_4329.pdf}.
-
-\bibitem{Famularo1999}[10.1016/S0005-1098(99)00058-8]
+\newblock \emph{\emph{IOSO optimization toolkit --- novel software to create better design}},
+\newblock in \emph{9th AIAA/ISSMO Symposium on Multidisciplinary Analysis and Optimization}, 2002. Available from \url{http://www.iosotech.com/text/2002\_4329.pdf}.
+
+\bibitem{Famularo1999} (MR1827451) [10.1016/S0005-1098(99)00058-8]
 \newblock D. Famularo, P. Pugliese and Y. D. Sergeyev,
-\newblock A global optimization technique for checking parametric robustness,
+\newblock \emph{\emph{A global optimization technique for checking parametric robustness}},
 \newblock \emph{Automatica}, \textbf{35} (1999), 1605--1611.
 
-\bibitem{fasanoPinter2013}[10.1007/978-1-4614-4469-5]
+\bibitem{fasanoPinter2013} (MR3013800) [10.1007/978-1-4614-4469-5]
 \newblock G. Fasano and J. D. Pint\'er,
-\newblock \emph{Modeling and optimization in space engineering},
+\newblock \emph{Modeling and Optimization in Space Engineering},
 \newblock Springer, 2013.
 
-\bibitem{floudasPardalosGOState}[10.1007/978-1-4613-3437-8]
+\bibitem{floudasPardalosGOState} (MR1390521) [10.1007/978-1-4613-3437-8]
 \newblock C. A. Floudas and M. P. Pardalos,
-\newblock \emph{State of the art in global optimization: computational methods and applications},
-\newblock Kluwer Academic Publishers, Dordrecht (1996).
-
-\bibitem{floudasPardGO}[10.1007/s10898-008-9332-8]
+\newblock \emph{State of the Art in Global Optimization: Computational Methods and Applications},
+\newblock Kluwer Academic Publishers, Dordrecht, 1996.
+
+\bibitem{floudasPardGO} (MR1147432) [10.1007/s10898-008-9332-8]
 \newblock C. A. Floudas and M. P. Pardalos,
-\newblock \emph{Recent advances in global optimization},
-\newblock Princeton University Press (2016).
-
-\bibitem{Gablonsky}[10.1023/A:1017930332101]
+\newblock \emph{Recent Advances in Global Optimization},
+\newblock Princeton University Press, 2016.
+
+\bibitem{Gablonsky} (MR1856800) [10.1023/A:1017930332101]
 \newblock J. M. Gablonsky and C. T. Kelley,
-\newblock A locally-biased form of the DIRECT algorithm,
-\newblock \emph{J. Glob. Optim.}, \textbf{21(1)} (2001), 27--37.
-
-\bibitem{gavianoKvasovLeraSergeev2003}[10.1145/962437.962444]
+\newblock \emph{\emph{A locally-biased form of the DIRECT algorithm}},
+\newblock \emph{J. Glob. Optim.}, \textbf{21} (2001), 27--37.
+
+\bibitem{gavianoKvasovLeraSergeev2003} (MR2077342) [10.1145/962437.962444]
 \newblock M. Gaviano, D. E. Kvasov, D. Lera and Y. D. Sergeev,
-\newblock Software for generation of classes of test functions with known local and global minima for global optimization,
-\newblock \emph{ACM Trans. Math. Software}, \textbf{29(4)} (2003), 469--480.
-
-\bibitem{gergelLebedev2015}[10.1016/j.procs.2015.11.008]
+\newblock \emph{\emph{Software for generation of classes of test functions with known local and global minima for global optimization}},
+\newblock \emph{ACM Trans. Math. Software}, \textbf{29} (2003), 469--480.
+
+\bibitem{gergelLebedev2015} [10.1016/j.procs.2015.11.008]
 \newblock V. Gergel and I. Lebedev,
-\newblock Heterogeneous parallel computations for solving global optimization problems,
+\newblock \emph{\emph{Heterogeneous parallel computations for solving global optimization problems}},
 \newblock \emph{Procedia Comput. Sci.}, \textbf{66} (2015), 53--62.
@@
 \newblock V. Gergel,
-\newblock A software system for multi-extremal optimization,
-\newblock \emph{Eur. J. Oper. Res.}, \textbf{65(3)} (1993), 305--313.
-
-\bibitem{gergel1996}
+\newblock \emph{\emph{A software system for multi-extremal optimization}},
+\newblock \emph{Eur. J. Oper. Res.}, \textbf{65} (1993), 305--313.
+
+\bibitem{gergel1996} (MR1399766)
 \newblock V. Gergel,
-\newblock A method for using derivatives in the minimization of multiextremum functions,
-\newblock \emph{Comput. Math. Math. Phys.}, \textbf{36(6)} (1996), 729--742.
-
-\bibitem{gergel1997}[10.1023/A:1008290629896]
+\newblock \emph{\emph{A method for using derivatives in the minimization of multiextremum functions}},
+\newblock \emph{Comput. Math. Math. Phys.}, \textbf{36} (1996), 729--742.
+
+\bibitem{gergel1997} (MR1443089) [10.1023/A:1008290629896]
 \newblock V. Gergel,
-\newblock A global optimization algorithm for multivariate functions with Lipschitzian first derivatives,
-\newblock \emph{J. Glob. Optim.}, \textbf{10(3)} (1997), 257--281.
-
-\bibitem{gergel2013}[doi:10.1016/j.procs.2013.05.164]
+\newblock \emph{\emph{A global optimization algorithm for multivariate functions with Lipschitzian first derivatives}},
+\newblock \emph{J. Glob. Optim.}, \textbf{10} (1997), 257--281.
+
+\bibitem{gergel2013} [doi:10.1016/j.procs.2013.05.164]
 \newblock V. Gergel, et al.,
-\newblock High performance computing in biomedical applications,
+\newblock \emph{\emph{High performance computing in biomedical applications}},
 \newblock \emph{Procedia Computer Science}, \textbf{18} (2013), 10--19.
 
-\bibitem{gergel2015}[10.15866/ireaco.v8i1.4935]
+\bibitem{gergel2015} [10.15866/ireaco.v8i1.4935]
 \newblock V. Gergel, et al.,
-\newblock Recognition of surface defects of cold-rolling sheets based on method of localities,
-\newblock \emph{International Review of Automatic Control }, \textbf{8(1)} (2015), 51--55.
-
-\bibitem{gergelSidorov2015}[10.1007/978-3-319-21909-7_49]
+\newblock \emph{\emph{Recognition of surface defects of cold-rolling sheets based on method of localities}},
+\newblock \emph{International Review of Automatic Control, } \textbf{8} (2015), 51--55.
+
+\bibitem{gergelSidorov2015} [10.1007/978-3-319-21909-7_49]
 \newblock V. Gergel and S. Sidorov,
-\newblock A two-level parallel global search algorithm for solving computationally intensive multi-extremal optimization problems,
+\newblock \emph{\emph{A two-level parallel global search algorithm for solving computationally intensive multi-extremal optimization problems}},
 \newblock \emph{LNCS}, \textbf{9251} (2015), 505--515.
 
-\bibitem{grishaginStrongin1984}
+\bibitem{grishaginStrongin1984} (MR785497)
 \newblock V. A. Grishagin and R. G. Strongin,
-\newblock Optimization of multi-extremal functions subject to monotonically unimodal constraints,
+\newblock \emph{\emph{Optimization of multi-extremal functions subject to monotonically unimodal constraints}},
 \newblock \emph{Engineering Cybernetics}, \textbf{5} (1984), 117--122.
 
-\bibitem{holmstromEdvall2004}[10.1007/978-1-4613-0215-5_19]
-\newblock K. Holmström and M. M. Edvall,
-\newblock The TOMLAB optimization environment,
-\newblock \emph{Modeling Languages in Mathematical Optimization},
-\newblock 369--376, Springer (2004).
-
-\bibitem{horstTuyGO}[10.1007/978-3-662-03199-5]
+\bibitem{holmstromEdvall2004} [10.1007/978-1-4613-0215-5_19]
+\newblock K. Holmstrm and M. M. Edvall,
+\newblock \emph{\emph{The TOMLAB optimization environment}},
+\newblock \emph{Modeling Languages in Mathematical Optimization}, Springer, (2004), 369--376.
+
+\bibitem{horstTuyGO} (MR1102239) [10.1007/978-3-662-03199-5]
 \newblock R. Horst and H. Tuy,
-\newblock \emph{Global optimization: deterministic approaches},
-\newblock Springer-Verlag, Berlin (1990).
-
-\bibitem{Jones}[10.1007/BF00941892]
+\newblock \emph{Global Optimization: Deterministic Approaches},
+\newblock Springer-Verlag, Berlin, 1990.
+
+\bibitem{Jones} (MR1246501) [10.1007/BF00941892]
 \newblock D. R. Jones, C. D. Perttunen and B. E. Stuckman,
-\newblock Lipschitzian optimization without the Lipschitz constant,
-\newblock \emph{J. Optim. Theory Appl.}, \textbf{79(1)} (1993), 157--181.
-
-\bibitem{kearfott2009}[10.1080/10556780802614051]
+\newblock \emph{\emph{Lipschitzian optimization without the Lipschitz constant}},
+\newblock \emph{J. Optim. Theory Appl.}, \textbf{79} (1993), 157--181.
+
+\bibitem{kearfott2009} (MR2554906) [10.1080/10556780802614051]
 \newblock R. B. Kearfott,
-\newblock GlobSol user guide,
+\newblock \emph{\emph{GlobSol user guide}},
 \newblock \emph{Optim. Methods Softw.}, \textbf{24} (2009), 687--708.
 
-\bibitem{Kvasov2015}[10.1016/j.advengsoft.2014.09.014]
+\bibitem{Kvasov2015} (MR3585540) [10.1016/j.advengsoft.2014.09.014]
 \newblock D. E. Kvasov and Y. D. Sergeyev,
-\newblock Deterministic approaches for solving practical black-box global optimization problems,
+\newblock \emph{\emph{Deterministic approaches for solving practical black-box global optimization problems}},
 \newblock \emph{Adv. Eng. Softw.}, \textbf{80} (2015), 58--66.
 
-\bibitem{Menniti}[10.1016/j.epsr.2007.10.009]
+\bibitem{Menniti} [10.1016/j.epsr.2007.10.009]
 \newblock D. E. Kvasov, D. Menniti, A. Pinnarelli, Y. D. Sergeyev and N. Sorrentino,
-\newblock Tuning fuzzy power-system stabilizers in multi-machine systems by global optimization algorithms based on efficient domain partitions,
-\newblock \emph{Electric Power Systems Research}, \textbf{78(7)} (2008), 1217--1229.
-
-\bibitem{Pizzuti}[10.1007/s00211-002-0419-8]
+\newblock \emph{\emph{Tuning fuzzy power-system stabilizers in multi-machine systems by global optimization algorithms based on
+ efficient domain partitions}},
+\newblock \emph{Electric Power Systems Research}, \textbf{78} (2008), 1217--1229.
+
+\bibitem{Pizzuti} (MR1971214) [10.1007/s00211-002-0419-8]
 \newblock D. E. Kvasov, C. Pizzuti and Y. D. Sergeyev,
-\newblock Local tuning and partition strategies for diagonal GO methods,
-\newblock \emph{Numerische Mathematik}, \textbf{94(1)} (2003), 93--106.
-
-\bibitem{liberti2009}[10.1007/0-387-30528-9_8]
+\newblock \emph{\emph{Local tuning and partition strategies for diagonal GO methods}},
+\newblock \emph{Numerische Mathematik}, \textbf{94} (2003), 93--106.
+
+\bibitem{liberti2009} (MR2206955) [10.1007/0-387-30528-9_8]
 \newblock L. Liberti,
-\newblock Writing global optimization software,
-\newblock in \emph{Nonconvex Optimization and Its Applications }, \textbf{84}, 211--262,
-\newblock Springer (2006).
-
-\bibitem{linSchrage2009}[10.1080/10556780902753221]
+\newblock \emph{\emph{Writing global optimization software}},
+\newblock in \emph{Nonconvex Optimization and Its Applications}, Springer, \textbf{84} (2006), 211--262.
+
+\bibitem{linSchrage2009} (MR2554904) [10.1080/10556780902753221]
 \newblock Y. Lin and L. Schrage,
-\newblock The global solver in the LINDO API,
-\newblock \emph{Optim. Methods Softw.}, \textbf{24(4-5)} (2009), 657--668.
-
-\bibitem{locatelliSchoenGO}[10.1137/1.9781611972672]
+\newblock \emph{\emph{The global solver in the LINDO API}},
+\newblock \emph{Optim. Methods Softw.}, \textbf{24} (2009), 657--668.
+
+\bibitem{locatelliSchoenGO} (MR3136805) [10.1137/1.9781611972672]
 \newblock M. Locatelli and F. Schoen,
-\newblock \emph{Global optimization: theory, algorithms and applications},
-\newblock SIAM (2013).
-
-\bibitem{luqueAlbaGA}[10.1007/978-3-642-22084-5]
+\newblock \emph{Global Optimization: Theory, Algorithms and Applications},
+\newblock SIAM, 2013.
+
+\bibitem{luqueAlbaGA} (MR2808873) [10.1007/978-3-642-22084-5]
 \newblock G. Luque and E. Alba,
-\newblock \emph{Parallel genetic algorithms. Theory and real world applications},
-\newblock Springer-Verlag, Berlin (2011).
-
-\bibitem{mongeauKarsentyRouze2000}[10.1080/10556780008805783]
-\newblock M. Mongeau, H. Karsenty, V. Rouzé and J. B. Hiriart-Urruty,
-\newblock Comparison of public-domain software for black box global optimization,
+\newblock \emph{Parallel Genetic Algorithms. Theory and Real World Applications},
+\newblock Springer-Verlag, Berlin, 2011.
+
+\bibitem{mongeauKarsentyRouze2000} (MR1785196) [10.1080/10556780008805783]
+\newblock M. Mongeau, H. Karsenty, V. Rouz$\acute{e}$ and J. B. Hiriart-Urruty,
+\newblock \emph{\emph{Comparison of public-domain software for black box global optimization}},
 \newblock \emph{Optim. Methods Softw.}, \textbf{13} (2000), 203--226.
 
-\bibitem{mullen2014}[10.18637/jss.v060.i06]
+\bibitem{mullen2014} [10.18637/jss.v060.i06]
 \newblock K. M. Mullen,
-\newblock Continuous global optimization in R,
-\newblock \emph{J. Stat. Softw.}, \textbf{60(6)} (2014).
-
-\bibitem{pardalosZhigljavskyZilinskas2016}[10.1007/978-3-319-29975-4]
+\newblock \emph{\emph{Continuous global optimization in R}},
+\newblock \emph{J. Stat. Softw.}, \textbf{60} (2014).
+
+\bibitem{pardalosZhigljavskyZilinskas2016} (MR2361744) [10.1007/978-3-319-29975-4]
 \newblock M. P. Pardalos, A. A. Zhigljavsky and J. \v Zilinskas,
-\newblock \emph{Advances in stochastic and deterministic global optimization},
-\newblock Springer (2016).
-
-\bibitem{pinterGO}[10.1007/978-1-4757-2502-5]
+\newblock \emph{Advances in Stochastic and Deterministic Global Optimization},
+\newblock Springer, 2016.
+
+\bibitem{pinterGO} (MR1374104) [10.1007/978-1-4757-2502-5]
 \newblock J. D. Pint\'er,
-\newblock \emph{Global optimization in action (Continuous and Lipschitz optimization: algorithms, implementations and applications)},
-\newblock Kluwer Academic Publishers, Dordrecht (1996).
-
-\bibitem{pinter2009}
+\newblock \emph{Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications)},
+\newblock Kluwer Academic Publishers, Dordrecht, 1996.
+
+\bibitem{pinter2009} (MR2528693)
 \newblock J. D. Pint\'er,
-\newblock Software development for global optimization,
+\newblock \emph{\emph{Software development for global optimization}},
 \newblock \emph{Lectures on Global Optimization. Fields Institute Communications}, \textbf{55} (2009), 183--204.
 
-\bibitem{riosSahinidis2013}[10.1007/s10898-012-9951-y]
+\bibitem{riosSahinidis2013} (MR3070154) [10.1007/s10898-012-9951-y]
 \newblock L. M. Rios and N. V. Sahinidis,
-\newblock Derivative-free optimization: a review of algorithms and comparison of software implementations,
+\newblock \emph{\emph{Derivative-free optimization: a review of algorithms and comparison of software implementations}},
 \newblock \emph{J. Glob. Optim.}, \textbf{56} (2013), 1247--1293.
 
-\bibitem{sahinidis1996}[10.1007/BF00138693]
+\bibitem{sahinidis1996} (MR1376505) [10.1007/BF00138693]
 \newblock N. V. Sahinidis,
-\newblock BARON: A general purpose global optimization software package,
+\newblock \emph{\emph{BARON: A general purpose global optimization software package}},
 \newblock \emph{J. Glob. Optim.}, \textbf{8} (1996), 201--205.
 
-\bibitem{sergeyev1995}[10.1137/0805041]
+\bibitem{sergeyev1995} (MR1358808) [10.1137/0805041]
 \newblock Y. D. Sergeyev,
-\newblock An information global optimization algorithm with local tuning,
-\newblock \emph{SIAM J. Optim.}, \textbf{5(4)} (1995), 858--870.
-
-\bibitem{sergeyev1999}
+\newblock \emph{\emph{An information global optimization algorithm with local tuning}},
+\newblock \emph{SIAM J. Optim.}, \textbf{5} (1995), 858--870.
+
+\bibitem{sergeyev1999} (MR1699326)
 \newblock Y. D. Sergeyev,
-\newblock Multidimensional global optimization using the first derivatives,
-\newblock \emph{Comput. Math. Math. Phys.}, \textbf{39(5)} (1999), 743--752.
-
-\bibitem{sergeyevKvasov2006}[	arXiv:1103.2056]
-\newblock Y. D. Sergeyev, D. E. Kvasov,
-\newblock Global search based on efficient diagonal partitions and a set of Lipschitz constants,
-\newblock \emph{SIAM Journal on Optimization}, \textbf{16(3)} (2006), 910--937.
-
-\bibitem{sergeyevGrishagin2001}[10.1023/A:1010185125012]
+\newblock \emph{\emph{Multidimensional global optimization using the first derivatives}},
+\newblock \emph{Comput. Math. Math. Phys.}, \textbf{39} (1999), 743--752.
+
+\bibitem{sergeyevKvasov2006} (MR2197562) [10.1137/040621132]
+\newblock Y. D. Sergeyev and D. E. Kvasov,
+\newblock \emph{\emph{Global search based on efficient diagonal partitions and a set of Lipschitz constants}},
+\newblock \emph{SIAM Journal on Optimization}, \textbf{16} (2006), 910--937.
+
+\bibitem{sergeyevGrishagin2001} (MR1825071) [10.1023/A:1010185125012]
 \newblock Y. D. Sergeyev and V. A. Grishagin,
-\newblock Parallel asynchronous global search and the nested optimization scheme,
-\newblock \emph{J. Comput. Anal. Appl.}, \textbf{3(2)} (2001), 123--145.
-
-\bibitem{sergeyevStronginLera2013}[10.1007/978-1-4614-8042-6]
+\newblock \emph{\emph{Parallel asynchronous global search and the nested optimization scheme}},
+\newblock \emph{J. Comput. Anal. Appl.}, \textbf{3} (2001), 123--145.
+
+\bibitem{sergeyevStronginLera2013} (MR3113120) [10.1007/978-1-4614-8042-6]
 \newblock Y. D. Sergeyev, R. G. Strongin and D. Lera,
-\newblock \emph{Introduction to global optimization exploiting space-filling curves},
-\newblock Springer (2013).
-
-\bibitem{Famularo2001}[10.1023/A:1012391611462]
+\newblock \emph{Introduction to Global Optimization Exploiting Space-filling Curves},
+\newblock Springer, 2013.
+
+\bibitem{Famularo2001} (MR1866703) [10.1023/A:1012391611462]
 \newblock Y. D. Sergeyev, D. Famularo and P. Pugliese,
-\newblock Index branch-and-bound algorithm for Lipschitz univariate global optimization with multiextremal constraints,
-\newblock \emph{J. Glob. Optim.}, \textbf{21(3)} (2001), 317--341.
-
-\bibitem{strongin1978}
+\newblock \emph{\emph{Index branch-and-bound algorithm for Lipschitz univariate global optimization with multiextremal constraints}},
+\newblock \emph{J. Glob. Optim.}, \textbf{21} (2001), 317--341.
+
+\bibitem{strongin1978} (MR509033)
 \newblock R. G. Strongin,
-\newblock \emph{Numerical methods in multi-extremal problems (information-statistical algorithms)},
-\newblock Moscow: Nauka, In Russian (1978).
-
-\bibitem{strongin1992}[10.1007/BF00122428]
+\newblock \emph{Numerical Methods in Multi-Extremal Problems (Information-Statistical Algorithms)},
+\newblock Moscow: Nauka, In Russian, 1978.
+
+\bibitem{strongin1992} (MR1263606) [10.1007/BF00122428]
 \newblock R. G. Strongin,
-\newblock Algorithms for multi-extremal mathematical programming problems employing a set of joint space-filling curves,
-\newblock \emph{J. Glob. Optim.}, \textbf{2(4)} (1992), 357--378.
+\newblock \emph{\emph{Algorithms for multi-extremal mathematical programming problems employing a set of joint space-filling curves}},
+\newblock \emph{J. Glob. Optim.}, \textbf{2} (1992), 357--378.
 
@@
 \newblock R. G. Strongin, V. P. Gergel, V. A. Grishagin and K. A. Barkalov,
-\newblock \emph{Parallel computations for global optimization problems},
-\newblock Moscow State University (In Russian), Moscow (2013).
-
-\bibitem{strSergGO}[10.1007/978-1-4615-4677-1]
+\newblock \emph{Parallel Computations for Global Optimization Problems},
+\newblock Moscow State University (In Russian), Moscow, 2013.
+
+\bibitem{strSergGO} (MR1797058) [10.1007/978-1-4615-4677-1]
 \newblock R. G. Strongin and Y. D. Sergeyev,
-\newblock \emph{Global optimization with non-convex constraints. Sequential and parallel algorithms},
+\newblock \emph{Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms},
 \newblock Kluwer Academic Publishers, Dordrecht (2000, 2nd ed. 2013, 3rd ed. 2014).
 
-\bibitem{zilinskTornGO}[10.1007/3-540-50871-6]
+\bibitem{zilinskTornGO} (MR1100586) [10.1007/3-540-50871-6]
 \newblock A. T\"orn and A. \v Zilinskas,
-\newblock \emph{Global optimization},
-\newblock Springer (1989).
+\newblock \emph{Global Optimization},
+\newblock Springer, 1989.
 
@@
 \newblock P. Venkataraman,
-\newblock \emph{Applied optimization with MATLAB programming},
-\newblock John Wiley \& Sons (2009).
-
-\bibitem{zhigljavskyRandGO}[10.1007/978-94-011-3436-1]
+\newblock \emph{Applied Optimization with MATLAB Programming},
+\newblock John Wiley \& Sons, 2009.
+
+\bibitem{zhigljavskyRandGO} (MR1187048) [10.1007/978-94-011-3436-1]
 \newblock A. A. Zhigljavsky,
-\newblock \emph{Theory of global random search},
-\newblock Kluwer Academic Publishers, Dordrecht (1991).
+\newblock \emph{Theory of Global Random Search},
+\newblock Kluwer Academic Publishers, Dordrecht, 1991.
 
@@
 \medskip
-% The data information below will be filled by AIMS editorial staff
-Received xxxx 20xx; revised xxxx 20xx.
+
+Received December 2016;  1$^{st}$ revision August 2017; final revision September  2017.
+
 \medskip

### 2018-03-28 2018/RSD/MultiEvolvents/RSD.tex
Заготовка статьи по множественным разверткам

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}
%%%%%
\usepackage{graphicx}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Множественные развертки\thanks{Выполнено при поддержке РНФ -...}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Alexander Sysoyev \and
Konstantin Barkalov \and
Vladislav Sovrasov \and
Ilya Lebedev \and
Victor Gergel \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
\email{alexander.sysoyev@itmm.unn.ru}\\
\email{konstantin.barkalov@itmm.unn.ru}\\
\email{sovrasov.vlad@gmail.com}\\
\email{ilya.lebedev@itmm.unn.ru}\\
\email{victor.gergel@itmm.unn.ru}\\
\url{http://hpc-education.unn.ru/основные-напраления/глобальная-оптимизация} 
\and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
15--250 words.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{First Section}
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}

### 2018-03-28 2018/RSD/MultiEvolvents/RSD.tex
Добавил название, аннотацию, разделы.

@@
 %
-\title{Множественные развертки\thanks{Выполнено при поддержке РНФ -...}}
+\title{Сравнение алгоритмов редукции размерности основанных на множественных отображений использующих криве Пеано\thanks{Выполнено при поддержке РНФ -...}}
 %
@@
 %
-\author{Alexander Sysoyev \and
-Konstantin Barkalov \and
+\author{Konstantin Barkalov \and
 Vladislav Sovrasov \and
-Ilya Lebedev \and
-Victor Gergel \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
+Ilya Lebedev \and}
 %
@@
 \email{victor.gergel@itmm.unn.ru}\\
-\url{http://hpc-education.unn.ru/основные-напраления/глобальная-оптимизация} 
-\and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
+\url{http://hpc-education.unn.ru/основные-напраления/глобальная-оптимизация} }
 %
@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-15--250 words.
+В работе рассматриваются параллельные алгоритмы решения задач многоэкстремальной оптимизации.
+Алгоритмы разработаны в рамках развиваемого в ННГУ им.
+Н.И.
+Лобачевского информационно-статистического подхода и реализованы в параллельной системе Globalizer.
+Решение задачи осуществляется путем редукции размерности исходной многомерной задачи к ряду информационно-связанных эквивалентных друг другу одномерных задач, которые решаются параллельно.
+Обсуждаются результаты вычислительных экспериментов, четырех типов отображений, произведенных на нескольких сотнях тестовых задач.
 
-\keywords{First keyword  \and Second keyword \and Another keyword.}
+\keywords{Global optimization \and  Dimension reduction \and Parallel algorithms  \and Multidimensional multiextremal optimization  \and Global search algorithms  \and Parallel computations }
 \end{abstract}
@@
 %
+
+\section{Statement of Multidimensional Global Optimization Problem}
+
+\section{Methods of Dimension Reduction}
+
+\subsection{Базовый алгоритм}
+s
+\subsection{Сдвиговые}
+s
+\subsection{Вращаемые}
+s
+\subsection{Еще одни развертки}
+s
+\subsection{Гладкие}
+s
+\section{Parallel Computations for Solving Global Optimization Problems.}
+\subsection{Core Multidimensional Algorithm of Global Search}
+s
+\subsection{Параллельные множественные отображения}
+s
+\section{Results of Numerical Experiments}
+s
+\subsection{Сравнение последовательных сдвиговых и вращаемых разверток}
+s
+\subsection{Параллельные вращаемые развертки}
+s
+\subsection{Параллельные вращаемые развертки для задач большой размерности}
+s
+\section{Вывод о целесообразности применения того или иного вида разверток для того или иного вида задач}
+s
+
 \section{First Section}

### 2018-04-04 2018/RSD/MultiEvolvents/RSD.tex
Добавил первую версию введения

@@
 %
-\title{Сравнение алгоритмов редукции размерности основанных на множественных отображений использующих криве Пеано\thanks{Выполнено при поддержке РНФ -...}}
+\title{Comparison of dimensionality reduction schemes for parallel global optimization algorithms\thanks{Выполнено при поддержке РНФ -...}}
 %
@@
 Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
-\email{alexander.sysoyev@itmm.unn.ru}\\
 \email{konstantin.barkalov@itmm.unn.ru}\\
@@
 \email{ilya.lebedev@itmm.unn.ru}\\
-\email{victor.gergel@itmm.unn.ru}\\
 \url{http://hpc-education.unn.ru/основные-напраления/глобальная-оптимизация} }
@@
 \begin{abstract}
-В работе рассматриваются параллельные алгоритмы решения задач многоэкстремальной оптимизации.
-Алгоритмы разработаны в рамках развиваемого в ННГУ им.
-Н.И.
-Лобачевского информационно-статистического подхода и реализованы в параллельной системе Globalizer.
-Решение задачи осуществляется путем редукции размерности исходной многомерной задачи к ряду информационно-связанных эквивалентных друг другу одномерных задач, которые решаются параллельно.
-Обсуждаются результаты вычислительных экспериментов, четырех типов отображений, произведенных на нескольких сотнях тестовых задач.
+This work considers a parallel algorithms for solving multi-extremal optimization problems.
+Algorithms are developed within the framework of the information-statistical approach and implemented in a parallel solver “Globalizer” .
+The optimization problem is solved by reducing the multidimensional problem to a set of joint one-dimensional problems that are solved in parallel.
+Five types of Peano-type space-filling curves are employed to reduce dimension.
+The results of computational experiments carried out on several hundred test problems are discussed.
 
@@
 %
+
+\section{Introduction}\label{sec:intro}
+
+Global (or multiextremal) optimization problems are among the most complex problems
+in both theory and practice of optimal decision making. In these kinds of problems,
+the optimization criterion can have several local optima within the search domain.
+The existence of several local optima makes
+finding the global optimum difficult essentially, since it requires examining the whole feasible
+search domain. The volume of computations for solving global optimization problems
+can increase exponentially with increasing number of varied parameters.
+\par
+These global optimization problem features impose special requirements on the quality
+of the optimization methods and on the software to implement these ones. The global
+optimization methods should be highly efficient, and the software systems should be developed on
+a good professional basis. In general, the global optimization problems can be solved at
+a reasonable time by employing parallel computations on modern supercomputing systems only.
+\par
+The general state of the art in the field of global optimization is presented in a
+number of key monographs  \cite{floudasPardGO}, \cite{horstTuyGO}, \cite{locatelliSchoenGO}, \cite{pinterGO}, \cite{strSergGO}, \cite{zilinskTornGO}, \cite{zhigljavskyRandGO}.
+The development of optimization methods, which use the high-performance computational
+systems to solve the time-consuming global optimization problems, is an area of
+intensive research --- see, for instance, \cite{censorZeniosParGO}, \cite{ciegisHentyParGO},
+\cite{luqueAlbaGA}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}. The obtained
+theoretical results provide the efficient solutions of many applied global
+optimization problems in various fields of scientific and technical applications \cite{Famularo1999}, \cite{fasanoPinter2013},
+\cite{floudasPardalosGOState}, \cite{Kvasov2015}, \cite{Menniti}, \cite{locatelliSchoenGO},
+\cite{luqueAlbaGA}, \cite{pardalosZhigljavskyZilinskas2016}, \cite{pinterGO}.
+\par
+At the same time, the practical implementation of these global optimization algorithms
+within the framework of industrial software systems is quite limited. In many cases,
+software implementations are experimental in nature and are used by the developers
+themselves to obtain the results from the computational experiments required for the
+scientific publications. This situation originates from high development costs of the
+professional software systems, which can be used by numerous users. In addition, the global
+optimization problems could be solved in an automatic mode rarely because of the
+complexity of these ones. The user should actively control the global search
+process that implies an adequate level of qualification in the field of optimization
+(particularly, the user should know and understand the global optimization methods well).
+\par
+In this work, the authors consider an approach to minimizing multiextremal functions  developed in ...????????
+This allows problems to be solved in which function values may not be deter-mined for the entire search domain.
+Under this approach, solving multidimen-sional problems is reduced (using Peano-type space-filling curves) to solving equivalent one-dimensional problems.
+It should be noted that standard approaches to algorithm parallelization are not quite applicable to global optimization.
+For example, the rules for selecting an-other iteration point are quite simple and do not require parallelization (as over-heads associated with organizing parallel computations will nullify any possible acceleration).
+Some acceleration can be achieved by parallelizing the computa-tion of function values describing the object to be optimized; however, this ap-proach is specific to each individual problem being solved.
+The following approach looks more promising.
+The algorithm can be modified to run several trials in parallel.
+This approach provides the efficiency (as paral-lelization is applied to the most computation-intensive part of the problem solv-ing process) and generality (in that it applies to a wide range of global optimiza-tion algorithms).
+The approach, described in  \cite{Two_Level_Parallel} for unconstrained optimization, was used in this work for parallelizing constrained optimization algorithms.
 
@@
 %
-\begin{thebibliography}{8}
+\begin{thebibliography}{107}
+
+\bibitem{barkalovGergel2014}
+\newblock K. Barkalov and V. Gergel,
+\newblock \emph{\emph{Multilevel scheme of dimensionality reduction for parallel global search algorithms}},
+\newblock in \emph{Proceedings of the 1st International Conference on Engineering and Applied Sciences Optimization}, (2014), 2111--2124.
+
+\bibitem{barkalovGergel2015}% (MR3539505) [10.1007/s10898-016-0411-y]
+\newblock K. Barkalov and V. Gergel,
+\newblock \emph{\emph{Parallel global optimization on GPU}},
+\newblock \emph{J. Glob. Optim.}, \textbf{66} (2016), 3--20.
+
+\bibitem{barkalovGergelLebedev2015} %[10.1007/978-3-319-21909-7_31]
+\newblock K. Barkalov, V. Gergel and I. Lebedev,
+\newblock \emph{\emph{Use of Xeon Phi coprocessor for solving global optimization problems}},
+\newblock \emph{LNCS}, \textbf{9251} (2015), 307--318.
+
+\bibitem{barkalovGergelLebedevSysoev2015}
+\newblock K. Barkalov, V. Gergel, I. Lebedev and A. Sysoev,
+\newblock \emph{\emph{Solving the global optimization problems on heterogeneous cluster systems}},
+\newblock in \emph{CEUR Workshop Proceedings}, \textbf{1482} (2015), 411--419.
+
+\bibitem{Barkalov2013}% [10.1007/978-3-642-39958-9_14]
+\newblock K. Barkalov, A. Polovinkin, I. Meyerov, S. Sidorov and N. Zolotykh,
+\newblock \emph{\emph{SVM regression parameters optimization using parallel global search algorithm}},
+\newblock \emph{LNCS}, \textbf{7979} (2013), 154--166.
+
+\bibitem{bussieckMeeraus2004} (MR3618583)% [10.1007/978-1-4613-0215-5_8]
+\newblock M. R. Bussieck and A. Meeraus,
+\newblock \emph{\emph{General algebraic modeling system (GAMS)}},
+\newblock in \emph{Modeling Languages in Mathematical Optimization} (ed. J. Kallrath), Springer, (2004), 137--157.
+
+\bibitem{censorZeniosParGO}% (MR1486040)
+\newblock Y. Censor and S. A. Zenios,
+\newblock \emph{Parallel Optimization: Theory, Algorithms, and Applications},
+\newblock Oxford University Press, 1998.
+
+\bibitem{ciegisHentyParGO} (MR2499546)% [10.1007/978-0-387-09707-7]
+\newblock R. \v Ciegis, D. Henty, B. K\aa gstr\"om and J. \v Zilinskas,
+\newblock \emph{Parallel Scientific Computing and Optimization: Advances and Applications},
+\newblock Springer, 2009.
+
+\bibitem{iosoDescription}
+\newblock I. N. Egorov, G. V. Kretinin, I. A. Leshchenko and S. V. Kuptzov,
+\newblock \emph{\emph{IOSO optimization toolkit --- novel software to create better design}},
+\newblock in \emph{9th AIAA/ISSMO Symposium on Multidisciplinary Analysis and Optimization}, 2002. Available from \url{http://www.iosotech.com/text/2002\_4329.pdf}.
+
+\bibitem{Famularo1999} %(MR1827451) [10.1016/S0005-1098(99)00058-8]
+\newblock D. Famularo, P. Pugliese and Y. D. Sergeyev,
+\newblock \emph{\emph{A global optimization technique for checking parametric robustness}},
+\newblock \emph{Automatica}, \textbf{35} (1999), 1605--1611.
+
+\bibitem{fasanoPinter2013}% (MR3013800) [10.1007/978-1-4614-4469-5]
+\newblock G. Fasano and J. D. Pint\'er,
+\newblock \emph{Modeling and Optimization in Space Engineering},
+\newblock Springer, 2013.
+
+\bibitem{floudasPardalosGOState} %(MR1390521) [10.1007/978-1-4613-3437-8]
+\newblock C. A. Floudas and M. P. Pardalos,
+\newblock \emph{State of the Art in Global Optimization: Computational Methods and Applications},
+\newblock Kluwer Academic Publishers, Dordrecht, 1996.
+
+\bibitem{floudasPardGO}% (MR1147432) [10.1007/s10898-008-9332-8]
+\newblock C. A. Floudas and M. P. Pardalos,
+\newblock \emph{Recent Advances in Global Optimization},
+\newblock Princeton University Press, 2016.
+
+\bibitem{Gablonsky} %(MR1856800) [10.1023/A:1017930332101]
+\newblock J. M. Gablonsky and C. T. Kelley,
+\newblock \emph{\emph{A locally-biased form of the DIRECT algorithm}},
+\newblock \emph{J. Glob. Optim.}, \textbf{21} (2001), 27--37.
+
+\bibitem{gavianoKvasovLeraSergeev2003} %(MR2077342) [10.1145/962437.962444]
+\newblock M. Gaviano, D. E. Kvasov, D. Lera and Y. D. Sergeev,
+\newblock \emph{\emph{Software for generation of classes of test functions with known local and global minima for global optimization}},
+\newblock \emph{ACM Trans. Math. Software}, \textbf{29} (2003), 469--480.
+
+\bibitem{gergelLebedev2015}% [10.1016/j.procs.2015.11.008]
+\newblock V. Gergel and I. Lebedev,
+\newblock \emph{\emph{Heterogeneous parallel computations for solving global optimization problems}},
+\newblock \emph{Procedia Comput. Sci.}, \textbf{66} (2015), 53--62.
+
+\bibitem{gergel1993}
+\newblock V. Gergel,
+\newblock \emph{\emph{A software system for multi-extremal optimization}},
+\newblock \emph{Eur. J. Oper. Res.}, \textbf{65} (1993), 305--313.
+
+\bibitem{gergel1996}% (MR1399766)
+\newblock V. Gergel,
+\newblock \emph{\emph{A method for using derivatives in the minimization of multiextremum functions}},
+\newblock \emph{Comput. Math. Math. Phys.}, \textbf{36} (1996), 729--742.
+
+\bibitem{gergel1997} %(MR1443089) [10.1023/A:1008290629896]
+\newblock V. Gergel,
+\newblock \emph{\emph{A global optimization algorithm for multivariate functions with Lipschitzian first derivatives}},
+\newblock \emph{J. Glob. Optim.}, \textbf{10} (1997), 257--281.
+
+\bibitem{gergel2013} %[doi:10.1016/j.procs.2013.05.164]
+\newblock V. Gergel, et al.,
+\newblock \emph{\emph{High performance computing in biomedical applications}},
+\newblock \emph{Procedia Computer Science}, \textbf{18} (2013), 10--19.
+
+\bibitem{gergel2015}% [10.15866/ireaco.v8i1.4935]
+\newblock V. Gergel, et al.,
+\newblock \emph{\emph{Recognition of surface defects of cold-rolling sheets based on method of localities}},
+\newblock \emph{International Review of Automatic Control, } \textbf{8} (2015), 51--55.
+
+\bibitem{gergelSidorov2015}% [10.1007/978-3-319-21909-7_49]
+\newblock V. Gergel and S. Sidorov,
+\newblock \emph{\emph{A two-level parallel global search algorithm for solving computationally intensive multi-extremal optimization problems}},
+\newblock \emph{LNCS}, \textbf{9251} (2015), 505--515.
+
+\bibitem{grishaginStrongin1984}% (MR785497)
+\newblock V. A. Grishagin and R. G. Strongin,
+\newblock \emph{\emph{Optimization of multi-extremal functions subject to monotonically unimodal constraints}},
+\newblock \emph{Engineering Cybernetics}, \textbf{5} (1984), 117--122.
+
+\bibitem{holmstromEdvall2004}% [10.1007/978-1-4613-0215-5_19]
+\newblock K. Holmstrm and M. M. Edvall,
+\newblock \emph{\emph{The TOMLAB optimization environment}},
+\newblock \emph{Modeling Languages in Mathematical Optimization}, Springer, (2004), 369--376.
+
+\bibitem{horstTuyGO}% (MR1102239) [10.1007/978-3-662-03199-5]
+\newblock R. Horst and H. Tuy,
+\newblock \emph{Global Optimization: Deterministic Approaches},
+\newblock Springer-Verlag, Berlin, 1990.
+
+\bibitem{Jones} (MR1246501)% [10.1007/BF00941892]
+\newblock D. R. Jones, C. D. Perttunen and B. E. Stuckman,
+\newblock \emph{\emph{Lipschitzian optimization without the Lipschitz constant}},
+\newblock \emph{J. Optim. Theory Appl.}, \textbf{79} (1993), 157--181.
+
+\bibitem{kearfott2009}%(MR2554906) [10.1080/10556780802614051]
+\newblock R. B. Kearfott,
+\newblock \emph{\emph{GlobSol user guide}},
+\newblock \emph{Optim. Methods Softw.}, \textbf{24} (2009), 687--708.
+
+\bibitem{Kvasov2015}% (MR3585540) [10.1016/j.advengsoft.2014.09.014]
+\newblock D. E. Kvasov and Y. D. Sergeyev,
+\newblock \emph{\emph{Deterministic approaches for solving practical black-box global optimization problems}},
+\newblock \emph{Adv. Eng. Softw.}, \textbf{80} (2015), 58--66.
+
+\bibitem{Menniti}% [10.1016/j.epsr.2007.10.009]
+\newblock D. E. Kvasov, D. Menniti, A. Pinnarelli, Y. D. Sergeyev and N. Sorrentino,
+\newblock \emph{\emph{Tuning fuzzy power-system stabilizers in multi-machine systems by global optimization algorithms based on
+ efficient domain partitions}},
+\newblock \emph{Electric Power Systems Research}, \textbf{78} (2008), 1217--1229.
+
+\bibitem{Pizzuti}% (MR1971214) [10.1007/s00211-002-0419-8]
+\newblock D. E. Kvasov, C. Pizzuti and Y. D. Sergeyev,
+\newblock \emph{\emph{Local tuning and partition strategies for diagonal GO methods}},
+\newblock \emph{Numerische Mathematik}, \textbf{94} (2003), 93--106.
+
+\bibitem{liberti2009}% (MR2206955) [10.1007/0-387-30528-9_8]
+\newblock L. Liberti,
+\newblock \emph{\emph{Writing global optimization software}},
+\newblock in \emph{Nonconvex Optimization and Its Applications}, Springer, \textbf{84} (2006), 211--262.
+
+\bibitem{linSchrage2009}% (MR2554904) [10.1080/10556780902753221]
+\newblock Y. Lin and L. Schrage,
+\newblock \emph{\emph{The global solver in the LINDO API}},
+\newblock \emph{Optim. Methods Softw.}, \textbf{24} (2009), 657--668.
+
+\bibitem{locatelliSchoenGO}% (MR3136805) [10.1137/1.9781611972672]
+\newblock M. Locatelli and F. Schoen,
+\newblock \emph{Global Optimization: Theory, Algorithms and Applications},
+\newblock SIAM, 2013.
+
+\bibitem{luqueAlbaGA}% (MR2808873) [10.1007/978-3-642-22084-5]
+\newblock G. Luque and E. Alba,
+\newblock \emph{Parallel Genetic Algorithms. Theory and Real World Applications},
+\newblock Springer-Verlag, Berlin, 2011.
+
+\bibitem{mongeauKarsentyRouze2000} %(MR1785196) [10.1080/10556780008805783]
+\newblock M. Mongeau, H. Karsenty, V. Rouz$\acute{e}$ and J. B. Hiriart-Urruty,
+\newblock \emph{\emph{Comparison of public-domain software for black box global optimization}},
+\newblock \emph{Optim. Methods Softw.}, \textbf{13} (2000), 203--226.
+
+\bibitem{mullen2014} %[10.18637/jss.v060.i06]
+\newblock K. M. Mullen,
+\newblock \emph{\emph{Continuous global optimization in R}},
+\newblock \emph{J. Stat. Softw.}, \textbf{60} (2014).
+
+\bibitem{pardalosZhigljavskyZilinskas2016}% (MR2361744) [10.1007/978-3-319-29975-4]
+\newblock M. P. Pardalos, A. A. Zhigljavsky and J. \v Zilinskas,
+\newblock \emph{Advances in Stochastic and Deterministic Global Optimization},
+\newblock Springer, 2016.
+
+\bibitem{pinterGO}% (MR1374104) [10.1007/978-1-4757-2502-5]
+\newblock J. D. Pint\'er,
+\newblock \emph{Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications)},
+\newblock Kluwer Academic Publishers, Dordrecht, 1996.
+
+\bibitem{pinter2009}% (MR2528693)
+\newblock J. D. Pint\'er,
+\newblock \emph{\emph{Software development for global optimization}},
+\newblock \emph{Lectures on Global Optimization. Fields Institute Communications}, \textbf{55} (2009), 183--204.
+
+\bibitem{riosSahinidis2013} %(MR3070154) [10.1007/s10898-012-9951-y]
+\newblock L. M. Rios and N. V. Sahinidis,
+\newblock \emph{\emph{Derivative-free optimization: a review of algorithms and comparison of software implementations}},
+\newblock \emph{J. Glob. Optim.}, \textbf{56} (2013), 1247--1293.
+
+\bibitem{sahinidis1996} %(MR1376505) [10.1007/BF00138693]
+\newblock N. V. Sahinidis,
+\newblock \emph{\emph{BARON: A general purpose global optimization software package}},
+\newblock \emph{J. Glob. Optim.}, \textbf{8} (1996), 201--205.
+
+\bibitem{sergeyev1995} %(MR1358808) [10.1137/0805041]
+\newblock Y. D. Sergeyev,
+\newblock \emph{\emph{An information global optimization algorithm with local tuning}},
+\newblock \emph{SIAM J. Optim.}, \textbf{5} (1995), 858--870.
+
+\bibitem{sergeyev1999}% (MR1699326)
+\newblock Y. D. Sergeyev,
+\newblock \emph{\emph{Multidimensional global optimization using the first derivatives}},
+\newblock \emph{Comput. Math. Math. Phys.}, \textbf{39} (1999), 743--752.
+
+\bibitem{sergeyevKvasov2006}% (MR2197562) [10.1137/040621132]
+\newblock Y. D. Sergeyev and D. E. Kvasov,
+\newblock \emph{\emph{Global search based on efficient diagonal partitions and a set of Lipschitz constants}},
+\newblock \emph{SIAM Journal on Optimization}, \textbf{16} (2006), 910--937.
+
+\bibitem{sergeyevGrishagin2001}% (MR1825071) [10.1023/A:1010185125012]
+\newblock Y. D. Sergeyev and V. A. Grishagin,
+\newblock \emph{\emph{Parallel asynchronous global search and the nested optimization scheme}},
+\newblock \emph{J. Comput. Anal. Appl.}, \textbf{3} (2001), 123--145.
+
+\bibitem{sergeyevStronginLera2013}% (MR3113120) [10.1007/978-1-4614-8042-6]
+\newblock Y. D. Sergeyev, R. G. Strongin and D. Lera,
+\newblock \emph{Introduction to Global Optimization Exploiting Space-filling Curves},
+\newblock Springer, 2013.
+
+\bibitem{Famularo2001}% (MR1866703) [10.1023/A:1012391611462]
+\newblock Y. D. Sergeyev, D. Famularo and P. Pugliese,
+\newblock \emph{\emph{Index branch-and-bound algorithm for Lipschitz univariate global optimization with multiextremal constraints}},
+\newblock \emph{J. Glob. Optim.}, \textbf{21} (2001), 317--341.
+
+\bibitem{strongin1978}% (MR509033)
+\newblock R. G. Strongin,
+\newblock \emph{Numerical Methods in Multi-Extremal Problems (Information-Statistical Algorithms)},
+\newblock Moscow: Nauka, In Russian, 1978.
+
+\bibitem{strongin1992}% (MR1263606) [10.1007/BF00122428]
+\newblock R. G. Strongin,
+\newblock \emph{\emph{Algorithms for multi-extremal mathematical programming problems employing a set of joint space-filling curves}},
+\newblock \emph{J. Glob. Optim.}, \textbf{2} (1992), 357--378.
+
+\bibitem{stronginGergelBarkalovParGO}
+\newblock R. G. Strongin, V. P. Gergel, V. A. Grishagin and K. A. Barkalov,
+\newblock \emph{Parallel Computations for Global Optimization Problems},
+\newblock Moscow State University (In Russian), Moscow, 2013.
+
+\bibitem{strSergGO}5 (MR1797058) [10.1007/978-1-4615-4677-1]
+\newblock R. G. Strongin and Y. D. Sergeyev,
+\newblock \emph{Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms},
+\newblock Kluwer Academic Publishers, Dordrecht (2000, 2nd ed. 2013, 3rd ed. 2014).
+
+\bibitem{zilinskTornGO}% (MR1100586) [10.1007/3-540-50871-6]
+\newblock A. T\"orn and A. \v Zilinskas,
+\newblock \emph{Global Optimization},
+\newblock Springer, 1989.
+
+\bibitem{venkataraman2009}
+\newblock P. Venkataraman,
+\newblock \emph{Applied Optimization with MATLAB Programming},
+\newblock John Wiley \& Sons, 2009.
+
+\bibitem{zhigljavskyRandGO}% (MR1187048) [10.1007/978-94-011-3436-1]
+\newblock A. A. Zhigljavsky,
+\newblock \emph{Theory of Global Random Search},
+\newblock Kluwer Academic Publishers, Dordrecht, 1991.
+
+\bibitem{Two_Level_Parallel}
+Gergel, V., Sidorov, S.: A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems.
+In: Malyshkin, V.
+(Ed.) PaCT 2015, LNCS, vol.
+9251, pp.
+505-515.
+Springer, Heidelberg (2015)
+
 \bibitem{ref_article1}

### 2018-04-04 2018/RSD/MultiEvolvents/RSD.tex
Добавил введение, описание базового алгоритма, параллельных множественных разверток и базовых разверток.

@@
 \section{Statement of Multidimensional Global Optimization Problem}
+In this paper, the core class of optimization problems, which can be solved using
+Globalizer, is formulated. This class involves the multidimensional global
+optimization problems without constraints, which can be defined in the following way:
+\begin{equation}
+\label{eq:task}
+\begin{array}{cr}\\
+  \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
+  D=\{y\in \mathbf{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
+\end{array}
+\end{equation}
+with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function \(\varphi(y)\) satisfies the Lipschitz condition
+\begin{equation}
+\label{eq:lip}
+|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
+\end{equation}
+where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbf{R}^N\) space.
+\par
+Usually, the minimized function \(\varphi(y)\) is defined as a computational procedure,
+according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
+(let us further call such a calculation \textit{a trial}). It is supposed that this procedure
+is a time-consuming one. As a result, the overall time of solving the optimization
+problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
+It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
+is highly important, since an estimate of the global minimum can be constructed on the
+basis of a finite number of computed values of the optimized function only in this case .
+\par
+As it has been shown earlier by many researchers
+(see, for instance, \cite{floudasPardalosGOState}, \cite{horstTuyGO}, \cite{pinterGO}, \cite{strSergGO}),
+finding the numerical estimate of the global optimum implies constructing a coverage of
+the search domain \(D\). As a result, the computational costs of solving the global
+optimization problems are readily very high even for a small number of varied parameters
+(the dimensionality of the problem). A notable reduction in the volume of computations
+can be achieved when the coverage of the search domain is non-uniform, i. e. the series
+of trial points is only dense in a vicinity of the global optimum point. The construction
+of such a non-uniform coverage could be provided in an adaptive way, when the selection
+of the next trial points is determined by using the search information (the preceding
+trial points and the values of the minimized function at these points) obtained in the course
+of computations. This necessary condition complicates considerably the computational schemes
+of global optimization methods since it implies a complex analysis of a large amount
+of multidimensional search information. As a result, many optimization algorithms use
+various approaches to the dimensional reduction \cite{pinterGO}, \cite{sergeyevStronginLera2013}, \cite{strongin1978}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
 
 \section{Methods of Dimension Reduction}
+\subsection{Базовый алгоритм}
 
-\subsection{Базовый алгоритм}
-s
+Within the framework of the information-statistical global optimization theory,
+the Peano space-filling curves (or evolvents) \(y(x)\) mapping the interval \([0,1]\)
+onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
+reduction \cite{sergeyevStronginLera2013}, \cite{strongin1978}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
+\par
+As a result of the reduction, the initial multidimensional global optimization
+problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
+\begin{equation}
+\label{eq:oneDimTask}
+\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
+\end{equation}
+\par
+It is important to note that this dimensionality reduction scheme transforms the minimized
+Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
+function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.
+\begin{equation}
+\label{eq:holder}
+|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
+\end{equation}
+where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
+constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem (\ref{eq:task}).
+\par
+The algorithms for the numerical construction of the Peano curve approximations are
+given in \cite{strSergGO}. As an illustration, an approximation of the Peano curve
+for the third density level is shown in Figure \ref{fig:peanoC}. Figure \ref{fig:peanoC}
+demonstrates the movement order in a two-dimensional domain to construct the Peano
+curve approximation; the precision of the Peano curve approximation is determined by the
+density level used in the construction.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.85\textwidth]{pictures/peanoC.eps}
+    \caption{A Peano curve approximation for the third density level}
+    \label{fig:peanoC}
+\end{figure}
+
+\par
+The computational scheme obtained as a result of the dimensionality reduction consists of the following
+(see Figure \ref{fig:peanoCUsage}):
+\begin{itemize}
+  \item The optimization algorithm performs the minimization of the reduced one-dimensional
+  function \(\varphi(y(x))\) from (\ref{eq:oneDimTask}),
+  \item After determining the next trial point \(x\), a multidimensional image \(y\) is calculated by using the
+mapping \(y(x)\),
+  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the point \(y\in D\),
+  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
+\end{itemize}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.55\textwidth]{pictures/peanoCUsage.eps}
+    \caption{The computational scheme for obtaining the value of the reduced one-dimensional function \(\varphi(y(x))\)}
+    \label{fig:peanoCUsage}
+\end{figure}
+
 \subsection{Сдвиговые}
@@
 \subsection{Core Multidimensional Algorithm of Global Search}
-s
+
+The information-statistical theory of global optimization formulated in \cite{strongin1978}, \cite{strSergGO} has
+served as a basis for the development of a large number of efficient multiextremal optimization
+methods \cite{barkalovGergel2014}, \cite{gergel1996}, \cite{gergel1997}, \cite{grishaginStrongin1984}, \cite{Pizzuti}, \cite{sergeyev1995}, \cite{sergeyev1999}, \cite{sergeyevGrishagin2001}, \cite{sergeyevStronginLera2013}, \cite{Famularo2001}.
+\par
+The optimization methods applied in Globalizer are based on the MAGS method, which can be presented as follows --- see \cite{strongin1978}, \cite{strSergGO}.
+\par
+Let us introduce a simpler notation for the optimization problem being solved
+\begin{equation}
+\label{eq:oneDimFunc}
+f(x) = \varphi(y(x)):x\in [0,1].
+\end{equation}
+\par
+The initial iteration of the algorithm is performed at an arbitrary point \mbox{\(x^1\in(0,1)\)}.
+Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
+The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the following rules.
+\par
+\textit{Rule 1}. Renumber the points of the preceding trials by the lower indices in order of increasing value of coordinates
+\begin{equation}
+  \label{step1}
+0=x_0<x_1<...<x_{k+1}=1,
+\end{equation}
+the points \(x_0\), \(x_{k+1}\) were introduced additionally for the convenience of further
+explanation, the values of the minimized function \(z_0\), \(z_{k+1}\) at these points are undefined.
+\par
+\textit{Rule 2}. Compute the current estimate of the H{\"o}lder constant \(H\) from (\ref{eq:holder})
+\begin{equation} \label{step2}
+M=\max_{1 < i\leq k}\frac{|z_i-z_{i-1}|}{\rho_i}, \;
+m = \left\{
+   \begin{array}{lr}
+     rM, & M > 0,\\
+     1, & M = 0,
+   \end{array}
+  \right.
+\end{equation}
+as the maximum of the relative differences of the minimized function values on the
+set of previously executed trial points \(x_i,1\leq i\leq k\) from (\ref{step1}).
+Hereafter, \(\rho_i=(x_i-x_{i-1})^\frac{1}{N},1\leq i\leq k+1\). The
+constant \(r\), \(r>1\), is the reliability parameter of the algorithm.
+\par
+\textit{Rule 3}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq k+1\), where
+\[
+R(i)=2\rho_i-4\frac{z_i}{m},\quad i=1,
+\]
+\begin{equation} \label{step3}
+R(i)=\rho_i+\frac{(z_i-z_{i-1})^2}{m^2\rho_i}-2\frac{z_i+z_{i-1}}{m},\quad 1<i<k+1, \\
+\end{equation}
+\[
+R(i)=2\rho_{i}-4\frac{z_{i-1}}{m},\quad i=k+1.
+\]
+
+\par
+\textit{Rule 4}. Determine the interval with the maximum characteristic
+\begin{equation} \label{step4}
+R(t)=\max_{1\leq i \leq k+1}R(i).
+\end{equation}
+\par
+\textit{Rule 5}. Execute a new trial at the point \(x^{k+1}\) located within the interval
+with the maximum characteristic from (\ref{step4})
+\begin{equation} \label{step5}
+  x^{k+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{r|z_{t}-z_{t-1}|}{m}\right]^N,\; \textrm{ if } 1<t<k+1,
+\end{equation}
+\[
+  x^{k+1}=\frac{x_t+x_{t-1}}{2},\; \textrm{ if } t=1 \textrm{ or } t=k+1.
+\]
+
+\par
+The stopping condition, which terminated the trials, is defined by the inequality
+\begin{equation}
+  \label{eq:stop_1}
+\rho_t<\varepsilon
+\end{equation}
+for the interval with the maximum characteristic from (\ref{step4}) and \(\varepsilon >0\) is the predefined
+accuracy of the optimization problem solution. If the stopping condition is not satisfied,
+the index \(k\) is incremented by 1, and the new global optimization iteration is executed.
+\par
+In order to explain the algorithm presented above, let us note the following.
+The characteristics \(R(i), 1\leq i\leq k+1\), calculated according to (\ref{step3}) could
+be interpreted as some measures of importance of the intervals with respect to the
+expected location of the global minimum point. Thus, the rules (\ref{step4}) and (\ref{step5}) for selecting
+the interval of the next trial become more clear --- the point of every next
+global optimization iteration is selected within the interval, where the global minimum
+point can be found most likely.
+\par
+The convergence conditions of the described algorithm are given, for example, in \cite{strSergGO}.
+
 \subsection{Параллельные множественные отображения}
-s
+
+The reduction of the multidimensional problems to the one-dimensional ones using evolvents has such important properties as the continuity and preservation of boundedness of function divided differences.
+However, a partial loss of in-formation on the nearness of the points in the multidimensional space takes place since a point \mbox{\(x\in[0,1]\)} has only the left and the right neighbors while the corresponding point  \mbox{\(y(x)\in R^N\)} has the neighbors in \(2N\) directions.
+As a result, when using the mappings like Peano curve the images \(y',y''\), which are close to each other in the N-dimensional space can correspond to the preimages \(x',x''\), which can be far away from each other in the interval [0,1].
+This property results in the excess computations since several limit points \(x',x''\) of the trial sequence generated by the index method in the interval [0,1] can correspond to a single limit point y in the \(N\)-dimensional space.
+\par
+One of the possible ways to overcome this disadvantage consists in using the multiple mapping \(Y^S (x)={y^1 (x),…,y^S (x)}\) instead of single evolvent \(y(x)\).
+To construct the set \(Y^S (x)\) different approaches can be used.
+For example, in \cite{strSergGO} a scheme was implemented, according to which each evolvent \(y^i (x)\) from \(Y^S (x)\) is constructed as a result of shifting the original evolvent \(y^0 (x)\) along the main diagonal of the hypercube \(D\).
+The set of Peano curves thus constructed allows one to obtain \(y',y''\) from \(D\) for any close multidimensional images, which differ only in one coordinate, close preimages \(x',x''\) from the interval [0,1] for the evolvent \(y^s (x),1≤s≤S\).
+Using the multiple mapping allows solving initial problem (1) by parallel solving the problems
+\par
+
+\[
+\min\{\varphi(y^s(x)):x\in [0,1]\}, 1≤s≤S
+\]
+
+on a set of intervals [0,1] by the index method.
+Each one-dimensional problem is solved on a separate processor.
+The trial results at the point \(x^k\) obtained for the problem being solved by particular processor are interpreted as the results of the trials in the rest problems (in the corresponding points \(x^(k_1 ),…,x^(k_S ))\).
+In this approach, a trial at the point \(x^k \in [0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of operations.
+\par
+1. Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
+\par
+2. Inform the rest of processors about the start of the trial execution at the point\( y^k\) (the blocking of the point \(y^k\) ).
+\par
+3.
+Determine the preimages \(x{}^{k_s}  \in [0,1], 1≤s≤S\), of the point \(y^k\) and interpret the trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points \(x{}^{k_1} ,…,x{}^{k_s} \)
+\par
+4. Inform the rest of processors about the trial results at the point \(y^k\).
+\par
+
+The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the sequential algorithm (except the method of the trial execution).
+Each processor has its own copy of the software realizing the computations of the problem functions and the decision rule of the index algorithm.
+For the organization of the interactions among the processors, the queues are created on each processor, where the processors store the information on the executed iterations in the form of the tuples: the processor number \(s\), the trial point \(x{}^{k_s}\).
+\par
+The proposed parallelization scheme was implemented with the use of MPI technology.
+Main features of implementation consist in the following.
+A separate MPI-process is created for each of \(S\) one-dimensional problems being solved, usually, one process per one processor employed.
+Each process can use p threads, usually one thread per an accessible core.
+\par
+At every iteration of the method, the process with the index \(s,0≤ s< S\) performs p trials in parallel at the points \(x^(s+iS),0≤i<p\).
+At that, each process stores all \(S_p\) points, and an attribute indicating whether this point is blocked by another process or not is stored for each point.
+Let us remind that the point is blocked if the process starts the execution of a trial at this point.
+\par
+At every iteration of the algorithm, operating within the \(s\)-th process, determines the coordinates of p «its own» trial points.
+Then, the interchange of the coordinates of images of the trial points \(y^(s+iS),0≤i<p, 0≤ s< S\) is performed (from each process to each one).
+After that, the preimages \(x^(q+iS),0≤ q<S,q≠s\) of the points received by the \(s\)-th process from the neighbor ones are determined with the use of the evolvent \(y^s (x)\).
+The points blocked within the \(s\)-th process will correspond to the preimages obtained.
+Then, each process performs the trials at the non-blocked points, the computations are performed in parallel using OpenMP.
+The results of the executed trials (the index of the point, the computed values of the problem functions, and the attribute of unblocking of this point) are transferred to all rest processes.
+All the points are added to the search information database, and the transition to the next iteration is performed.
+\par
+
 \section{Results of Numerical Experiments}

### 2018-04-05 2018/RSD/MultiEvolvents/RSD.tex
RSD: add a sections about evolvents

@@
 %%%%%
+\usepackage{eqnarray,amsmath}
+\usepackage{subfig}
 \usepackage{graphicx}
-
+\usepackage{amsfonts}
 % Used for displaying a sample figure. If possible, figure files should
@@
 %
-\title{Comparison of dimensionality reduction schemes for parallel global optimization algorithms\thanks{Выполнено при поддержке РНФ -...}}
+\title{Comparison of dimensionality reduction schemes for parallel global optimization algorithms
+\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150}}
 %
@@
 Vladislav Sovrasov \and
-Ilya Lebedev \and}
+Ilya Lebedev}
 %
-\authorrunning{F. Author et al.}
+\authorrunning{K. Barkalov et al.}
 % First names are abbreviated in the running head.
@@
 \subsection{Сдвиговые}
-s
+\label{sec:shifted}
+
+The reduction of the multidimensional problems to the one-dimensional ones using the Peano curves has such important properties as the continuity and the uniform boundedness of the function differences for limited variation of argument.
+However, a partial loss of information on the nearness of the points in the multidimensional space takes place since a point $x \in [0,1]$ has the left and the right neighbors only while the corresponding point $y(x) \in D \subset R^N$ has the neighbors in $2N$ directions.
+As a result, when using the mappings like Peano curve the images $y',\ y'',$ which are close to each other in the $N$-dimensional space can correspond to the preimages $x',\ x'',$ which can be far away from each other in the interval $[0,1]$.
+This property results in the excess computations since several limit points $x',\ x''$ of the trial sequence generated by the index method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional space.
+
+One of the possible ways to overcome this disadvantage consists in using the multiple mappings
+\begin{equation}%\label{eq:142}
+Y_L(x)=\left\{y^0(x),\ y^1(x),...,\ y^L(x)\right\}
+\end{equation}
+instead of single Peano curve $y(x)$ (see \cite{Strongin1991,Strongin1992,strSergGO}).
+
+Let us consider a family of the hypercubes
+\begin{equation}\label{6_hypercubes}
+D_l= \left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq L,
+\end{equation}
+where the hypercube $D_{l+1}$ is obtained by the translation of the hypercube $D_l$ along the main diagonal with displacement equal to $2^{-l}$ in each coordinate.
+In Fig.~\ref{fig:shifted_ev} the hypercubes $D_0,...,D_3$ for the case $N=2,\ L=3$ are presented.
+
+\begin{figure}[ht]
+    \centering
+    \subfloat[Two mappings]{{\includegraphics[width=.6\textwidth]{pictures/shifted.pdf}}\label{fig:shifted_ev}}
+    \subfloat[Hypercubes $D_l$]{{\includegraphics[width=.4\textwidth]{pictures/shifted_cube.png}}\label{fig:shifted_cube}}
+    \caption{Multiple mappings}
+\end{figure}
+
+Let us assume that the evolvent $y^0(x)$  maps the interval $[0,1]$ onto the hypercube $D_0$ from (\ref{6_hypercubes}), i.e.,
+\[
+D_0 = \left\{y^0(x) : x \in [0,1]\right\}.
+\]
+Then, the evolvents $y^l(x)=\left\{y_1^l(x),...,y_N^l(x)\right\}$, the coordinates of which are defined by the conditions
+\[
+y_i^l(x)=y_i^{l-1}(x)+2^{-l},\ 1\leq i\leq N, \ 1\leq l\leq L,
+\]
+map the interval $[0,1]$ onto the corresponding hypercubes $D_l,\ 1\leq l \leq L$.
+In Fig.~\ref{fig:shifted_ev} the image of the interval $[0,1]$ obtained by the curve $y^0(x),\ x\in [0,1],$ is shown as the dashed line.
+Since the hypercube $D$ from (\ref{eq:task}) is included in the common part of the family of hypercubes (\ref{6_hypercubes}) (the boundaries of hypercube $D$ are highlighted in Fig.~\ref{fig:shifted_cube}), having introduced an additional constraint function
+\begin{equation}\label{6_g0}
+g_0(y)=\max\left\{\left|y_i\right| - 2^{-1}:\ 1\leq i\leq N\right\},
+\end{equation}
+one can present the initial hypercube $D$ in the form
+\[
+D=\left\{y^l(x):\; x\in [0,1],\ g_0(y^l(x))\leq 0 \right\},\ 0\leq l \leq L,
+\]
+i.e., $g_0(y) \leq 0$ if $y\in D$ and $g_0(y)>0$ otherwise. Consequently, any point $y \in D$ has its own preimage $x^l \in [0,1]$ for each mapping $y^l(x),\ 0\leq l\leq L$.
+
+Thus, each evolvent $y^l(x),\ 0\leq l \leq L,$ generates its own problem of the type (\ref{eq:task}) featured by its own extended (in comparison with $D$) search domain $D_l$ and the additional constraint with the left hand part from (\ref{6_g0})
+\begin{equation}\label{6_problem_l}
+\min{\left\{\varphi(y^l(x)):x\in [0,1], \; g_j(y^l(x))\leq 0, \; 0 \leq j \leq m\right\}}, \ 0 \leq l \leq L.
+\end{equation}
+
+The problems (\ref{6_problem_l}) correspond to the domains $Q_0^l=[0,1]$ and $Q_{j+1}^l, 0 \leq j \leq m,$ defined by the expression
+\[
+Q_{j+1}^l = \left\{x \in Q_j^l:g_j(y^l(x))\leq 0\right\},\ 0\leq j\leq m,
+\]
+for the corresponding evolvents $y^l(x)$. The application of a multiple mapping defines the following relation for the nearness in the multidimensional search domain and in the one-dimensional one.
+
+\begin{theorem}
+Let a point $y^\ast$ from the domain $D$ be contained in the line segment with the end-points $y',y'' \in D$ meeting the requirements
+\[
+\left|y'_j - y''_j\right|\leq 2^{-p},\ y'_i=y''_i=y^\ast_i, \ 1\leq i \leq N,\ i \neq j,
+\]
+where $p$ is an integer  and $1\leq p \leq L$, i.e., the line segment is collinear with the $j$-th axis in $R^N$.
+Then, there exist at least one mapping $y^l(x),\ 0\leq l\leq L$, and the preimages $x^\ast,\; x',\; x''\in [0,1]$ such that
+\[
+y^\ast = y^l(x^\ast),\ y'=y^l(x'),\ y'' = y^l(x'')
+\]
+and
+\[
+\max \left\{ \left|x'-x^\ast\right|,\; \left|x''-x^\ast\right|,\; \left|x'-x''\right|\right\} \leq 2^{-pN}.
+\]
+\end{theorem}
+\begin{proof}
+Proof of this theorem is given in \cite{strSergGO}.
+\qed
+\end{proof}
+
+The conditions of the theorem single out a specific vicinity of the point $y^\ast$.
+This vicinity includes only the points, which can be obtained by the shift of $y^\ast$ parallel to one of the coordinate axes with a displacement not more than $2^{-p}$.
+By changing  $j,\ 1\leq j\leq N,$ in the theorem conditions it is possible to obtain the neighbours in any $N$ coordinate directions.
+According to the statement, the closeness of the points in the $N$-dimensional space in a particular direction will be reflected by the closeness of their preimages in one of the univariate problems.
+The information on the closeness of the points results, first, in more precise estimate of Lipschitz constants and, second, in the increase of the characteristics of the intervals, the images of the end points of which are close to each other in the $N$-dimensional space.
+
 \subsection{Вращаемые}
-s
-\subsection{Еще одни развертки}
-s
-\subsection{Гладкие}
-s
+The application of the scheme for building the multiple evolvents (hereinafter called the shifted evolvents or $S$-evolvents) described in Subsection \ref{sec:shifted} allows to preserve the information on the nearness of the points in the multidimensional space and, therefore, to provide more precise (as compared to a single evolvent) estimate of Lipschitz constant in the search process.
+However, this approach has serious restrictions, which narrow the applicability of the parallel algorithms, designed on the base of the $S$-evolvents.
+
+Because a shifted evolvent is built as a result of the shift of a hypercube $D$ along the main diagonal, and the shift step decreases 2 times for the building of each subsequent mapping, the number of such shifts is limited by the evolvent density $m$.
+In the case when the shift step is less than $2^{-m}$, the next evolvent coincides with the previous one.
+Thus, the applicable number of the evolvents $L$ and, hence, the number of processors are limited by $m$ ($L\leq m$), where $m$ is the density of the evolvents.
+
+Also, it follows from the algorithm for building the shifted evolvents that solving the initial problem in the domain $D$ is reduced to solving a set of problems in the extended domains $D_l$ from (\ref{6_hypercubes}) with additional constraint (\ref{6_g0}).
+As a result, the structure of the search domain becomes more complicated, and the exponential decrease of the volume of the search domain $D$ in relation to the volumes of the domains $D_l$ with increasing dimensionality of the problem takes place.
+
+To overcome the disadvantages mentioned above and to preserve the information on the nearness of the points in the $N$-dimensional space, a novel scheme of building of the multiple mappings is proposed.
+The building of a set of Peano curves not by the shift along the main diagonal of the hypercube but by rotation of the evolvents around the coordinate origin is a distinctive feature of the proposed scheme \cite{Gergel2009}.
+In the initial non-rotated mapping for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other.
+In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer.
+The evolvents generated according to the novel scheme will be hereinafter called the rotated evolvents or $R$-evolvents.
+
+In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case $N=2$ are presented as an illustration.
+The maximum number of various rotations of the evolvents mapping the $N$-dimensional hypercube onto a one-dimensional interval is $2^N$.
+The employment of all possible rotations might appear to be redundant.
+In this case, one can select only a part of all rotations.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.8\linewidth]{pictures/rotated_1.pdf}
+  \caption{Two rotated evolvents on the same plane}
+  \label{6_fig_9}
+\end{figure}
+
+As a possible approach, one can propose to generate the new evolvents by the rotation of the initial evolvents on the angle of $\pm\pi/2$ in each of the coordinate planes.
+For example, the rotation matrices in the plane $(y_2,y_4)$ for  $N=5$ are presented below.
+\begin{eqnarray*}
+ \begin{pmatrix}
+  1 & 0 & 0 & 0 & 0 \\
+  0 & 0 & 0 & -1 & 0 \\
+  0 & 0 & 1 & 0 & 0 \\
+  0 & 1 & 0 & 0 & 0 \\
+  0 & 0 & 0 & 0 & 1
+ \end{pmatrix}
+\qquad \qquad
+ \begin{pmatrix}
+  1 & 0 & 0 & 0 & 0 \\
+  0 & 0 & 0 & 1 & 0 \\
+  0 & 0 & 1 & 0 & 0 \\
+  0 & -1 & 0 & 0 & 0 \\
+  0 & 0 & 0 & 0 & 1
+ \end{pmatrix}
+\end{eqnarray*}
+
+The number of such pairs of rotations is determined by the number of the coordinate planes in the space, which equals to  $N(N-1)/2$.
+The total number of the transformations equals to $N(N-1)$.
+Taking into account the initial mapping, one can conclude that this method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals.
+Moreover, the additional constraint  $g_0(y) \leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent.
+This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.
+
+The use of the set of mappings $Y_L(x)=\{y^1(x),...,y^L(x)\}$ results in the appearing of the corresponding set of the one-dimensional multiextremal problems
+\begin{equation}\label{6_problem_lr}
+  \min{\left\{\varphi(y^l(x)):x\in [0,1], \; g_j(y^l(x))\leq 0, \; 1 \leq j \leq m\right\}}, \ 1 \leq l \leq L.
+\end{equation}
+Each problem from this set can be solved independently.
+Any computation of the value $z=g_\nu(y'),\ y'=y^i(x')$ of the function $g_\nu(y)$ in the $i$-th problem can be interpreted as a computation of the value $z=g_\nu(y'),\ y'=y^s(x'')$ for any other $s$-th problem without time-comsuming computations of the functions $g_\nu(y)$.
+Such information integrity allows to solve  initial problem (\ref{eq:task}) by solving $L$ problems (\ref{6_problem_lr}) on a set of intervals $[0,1]$ in parallel by the index method.
+
+\subsection{Non injective evolvent}
+
+Как уже было сказано в секции \ref{sec:shifted}, потеря информации о близости точек в многомерном пространстве может быть частично скомпенсирована использованием множественных отображений $Y_L(x)=\{y^1(x),...,y^L(x)\}$.
+Однако, сама по себе кривая типа Пеано сохраняет в себе часть этой информации: она не является инъективым отображением, поэтому имея один образ $y(x)\in \mathbb{R}^N$, можно получить несколько несколько отличных $x$ прообразов $t_j\in[0,1], t_j \not = x$, которые затем могут быть добавлены в поисковую информацию индексного метода.
+
+Кривая типа Пеано, используемая в (\ref{eq:oneDimTask}) для редукции размерности, определяется через предельный переход,
+поэтому не может быть вычислена непосредтвенно. При численной оптимизации используется некоторое её приближение, являющееся
+инъективной кусочно-линейной кривой.
+В \cite{strongin1978} было предложено неинъективное отображение сетки на отрезка $[0,1]$ на сетку в гиперкубе $D$.
+Каждый многомерный узел может иметь до $2^N$ одномерных прообразов.
+На рис.
+\ref{fig:noninjective} изображена кривая с самопересечениями, полученная при соединении узлов грубой многомерной сетки в порядке следования их прообразов из отрезка $2^N$, а также отмечана точка, имеющая 3 прообраза.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.6\linewidth]{pictures/noninjective.pdf}
+  \caption{Numerical non injective evolvent}
+  \label{fig:noninjective}
+\end{figure}
+
+Недостатком неинъективной развёртки является потенциально большое количетво прообразов (до $2^N$) и невозможность использования параллельной схемы для множественных отображений из секции \ref{sec:parallel_evolvents}.
+
+\subsection{Smooth evolvent}
+
 \section{Parallel Computations for Solving Global Optimization Problems.}
@@
 \subsection{Параллельные множественные отображения}
-
+\label{sec:parallel_evolvents}
 The reduction of the multidimensional problems to the one-dimensional ones using evolvents has such important properties as the continuity and preservation of boundedness of function divided differences.
@@
 
-\section{First Section}
-\subsection{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is
-not indented. The first paragraph that follows a table, figure,
-equation etc. does not need an indent, either.
-
-Subsequent paragraphs, however, are indented.
-
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig1}
-\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-%
 % ---- Bibliography ----
@@
 
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
+\bibitem{Strongin1991}
+Strongin, R.G.: Parallel multi-extremal optimization using a set of evolvents. Comp. Math. Math. Phys. \textbf{31(8)}, 37--46 (1991)
 
-\bibitem{ref_lncs1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
+\bibitem{Strongin1992}
+Strongin, R.G.: Algorithms for multi-extremal mathematical programming problems employing the set of joint space-filling curves. J. Global Optim. \textbf{2(4)}, 357--378 (1992)
 
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
+\bibitem{Gergel2009}
+Strongin, R.G., Gergel, V.P., Barkalov, K.A.: Parallel methods for global optimization problem solving. Journal of instrument engineering. \textbf{52}, 25--33 (2009) (In Russian)
 \end{thebibliography}

### 2018-04-07 2018/RSD/MultiEvolvents/RSD.tex
RSD: add new experiments

@@
 \usepackage{amsfonts}
+\usepackage{amssymb}
 % Used for displaying a sample figure. If possible, figure files should
@@
 In this work, the authors consider an approach to minimizing multiextremal functions developed in ...????????
-This allows problems to be solved in which function values may not be deter-mined for the entire search domain.
+This allows problems to be solved in which function values may not be determined for the entire search domain.
 Under this approach, solving multidimen-sional problems is reduced (using Peano-type space-filling curves) to solving equivalent one-dimensional problems.
@@
 \par
-The computational scheme obtained as a result of the dimensionality reduction consists of the following
-(see Figure \ref{fig:peanoCUsage}):
+The computational scheme obtained as a result of the dimensionality reduction consists of the following:
+%(see Figure \ref{fig:peanoCUsage}):
 \begin{itemize}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.55\textwidth]{pictures/peanoCUsage.eps}
-    \caption{The computational scheme for obtaining the value of the reduced one-dimensional function \(\varphi(y(x))\)}
-    \label{fig:peanoCUsage}
-\end{figure}
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=0.55\textwidth]{pictures/peanoCUsage.eps}
+%    \caption{The computational scheme for obtaining the value of the reduced one-dimensional function \(\varphi(y(x))\)}
+%    \label{fig:peanoCUsage}
+%\end{figure}
 
@@
 In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case $N=2$ are presented as an illustration.
-The maximum number of various rotations of the evolvents mapping the $N$-dimensional hypercube onto a one-dimensional interval is $2^N$.
-The employment of all possible rotations might appear to be redundant.
-In this case, one can select only a part of all rotations.
 
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{pictures/rotated_1.pdf}
+  \includegraphics[width=0.6\linewidth]{pictures/rotated_1.pdf}
   \caption{Two rotated evolvents on the same plane}
@@
 
-As a possible approach, one can propose to generate the new evolvents by the rotation of the initial evolvents on the angle of $\pm\pi/2$ in each of the coordinate planes.
-For example, the rotation matrices in the plane $(y_2,y_4)$ for  $N=5$ are presented below.
-\begin{eqnarray*}
- \begin{pmatrix}
-  1 & 0 & 0 & 0 & 0 \\
-  0 & 0 & 0 & -1 & 0 \\
-  0 & 0 & 1 & 0 & 0 \\
-  0 & 1 & 0 & 0 & 0 \\
-  0 & 0 & 0 & 0 & 1
- \end{pmatrix}
-\qquad \qquad
- \begin{pmatrix}
-  1 & 0 & 0 & 0 & 0 \\
-  0 & 0 & 0 & 1 & 0 \\
-  0 & 0 & 1 & 0 & 0 \\
-  0 & -1 & 0 & 0 & 0 \\
-  0 & 0 & 0 & 0 & 1
- \end{pmatrix}
-\end{eqnarray*}
-
-The number of such pairs of rotations is determined by the number of the coordinate planes in the space, which equals to  $N(N-1)/2$.
-The total number of the transformations equals to $N(N-1)$.
-Taking into account the initial mapping, one can conclude that this method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals.
+Taking into account the initial mapping, one can conclude that current implementation of the method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals.
 Moreover, the additional constraint  $g_0(y) \leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent.
@@
 инъективной кусочно-линейной кривой.
-В \cite{strongin1978} было предложено неинъективное отображение сетки на отрезка $[0,1]$ на сетку в гиперкубе $D$.
+В \cite{strongin1978} было предложено неинъективное отображение равномерной сетки на отрезка $[0,1]$ на равномерную сетку в гиперкубе $D$.
 Каждый многомерный узел может иметь до $2^N$ одномерных прообразов.
 На рис.
-\ref{fig:noninjective} изображена кривая с самопересечениями, полученная при соединении узлов грубой многомерной сетки в порядке следования их прообразов из отрезка $2^N$, а также отмечана точка, имеющая 3 прообраза.
+\ref{fig:noninjective} изображена кривая с самопересечениями, полученная при соединении узлов грубой многомерной сетки в порядке следования их прообразов из отрезка $[0,1]$, а также отмечана точка, имеющая 3 прообраза.
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.6\linewidth]{pictures/noninjective.pdf}
-  \caption{Numerical non injective evolvent}
-  \label{fig:noninjective}
-\end{figure}
+%\begin{figure}[t]
+%  \centering
+%  \includegraphics[width=0.6\linewidth]{pictures/noninjective.pdf}
+%  \caption{Numerical non injective evolvent}
+%  \label{fig:noninjective}
+%\end{figure}
 
@@
 \subsection{Smooth evolvent}
+
+%\begin{figure}[t]
+%  \centering
+%  \includegraphics[width=0.6\linewidth]{pictures/smooth.pdf}
+%  \caption{Smooth evolvent}
+%  \label{fig:smooth}
+%\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \subfloat[Smooth evolvent]{{\includegraphics[width=.6\textwidth]{pictures/smooth.pdf}}\label{fig:smooth}}
+    \subfloat[Numerical non injective evolvent]{{\includegraphics[width=.4\textwidth]{pictures/noninjective.pdf}}\label{fig:noninjective}}
+    \caption{Different evolvents}
+\end{figure}
 
@@
 \label{sec:parallel_evolvents}
-The reduction of the multidimensional problems to the one-dimensional ones using evolvents has such important properties as the continuity and preservation of boundedness of function divided differences.
-However, a partial loss of in-formation on the nearness of the points in the multidimensional space takes place since a point \mbox{\(x\in[0,1]\)} has only the left and the right neighbors while the corresponding point  \mbox{\(y(x)\in R^N\)} has the neighbors in \(2N\) directions.
-As a result, when using the mappings like Peano curve the images \(y',y''\), which are close to each other in the N-dimensional space can correspond to the preimages \(x',x''\), which can be far away from each other in the interval [0,1].
-This property results in the excess computations since several limit points \(x',x''\) of the trial sequence generated by the index method in the interval [0,1] can correspond to a single limit point y in the \(N\)-dimensional space.
-\par
-One of the possible ways to overcome this disadvantage consists in using the multiple mapping \(Y^S (x)={y^1 (x),…,y^S (x)}\) instead of single evolvent \(y(x)\).
-To construct the set \(Y^S (x)\) different approaches can be used.
-For example, in \cite{strSergGO} a scheme was implemented, according to which each evolvent \(y^i (x)\) from \(Y^S (x)\) is constructed as a result of shifting the original evolvent \(y^0 (x)\) along the main diagonal of the hypercube \(D\).
-The set of Peano curves thus constructed allows one to obtain \(y',y''\) from \(D\) for any close multidimensional images, which differ only in one coordinate, close preimages \(x',x''\) from the interval [0,1] for the evolvent \(y^s (x),1≤s≤S\).
 Using the multiple mapping allows solving initial problem (1) by parallel solving the problems
-\par
-
 \[
@@
 \]
-
 on a set of intervals [0,1] by the index method.
@@
 \par
-At every iteration of the method, the process with the index \(s,0≤ s< S\) performs p trials in parallel at the points \(x^(s+iS),0≤i<p\).
+At every iteration of the method, the process with the index \(s,0\leqslant s< S\) performs p trials in parallel at the points \(x^{(s+i_S)},0\leqslant i<p\).
 At that, each process stores all \(S_p\) points, and an attribute indicating whether this point is blocked by another process or not is stored for each point.
@@
 At every iteration of the algorithm, operating within the \(s\)-th process, determines the coordinates of p «its own» trial points.
-Then, the interchange of the coordinates of images of the trial points \(y^(s+iS),0≤i<p, 0≤ s< S\) is performed (from each process to each one).
-After that, the preimages \(x^(q+iS),0≤ q<S,q≠s\) of the points received by the \(s\)-th process from the neighbor ones are determined with the use of the evolvent \(y^s (x)\).
+Then, the interchange of the coordinates of images of the trial points \(y^{(s+i_S)},0\leqslant i<p, 0\leqslant s< S\) is performed (from each process to each one).
+After that, the preimages \(x^{(q+i_S)},0\leqslant q<S,q\not=s\) of the points received by the \(s\)-th process from the neighbor ones are determined with the use of the evolvent \(y^s (x)\).
 The points blocked within the \(s\)-th process will correspond to the preimages obtained.
@@
 \section{Results of Numerical Experiments}
-s
-\subsection{Сравнение последовательных сдвиговых и вращаемых разверток}
-s
+The computational experiments have been carried out on the Lobachevsky supercomputer at
+State University of Nizhny Novgorod. A computational node included 2 Intel
+Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. The CPUs had 8 cores (i. e. total 16 cores
+were available per a node). Все рассматриваемые алгортимы и развёртки были реализованы на языке C++ в рамках программной системы Globalizer.
+Для реализации параллелизма на одном узле использована технология OpenMP, а для параллелизма на несколько узлов --- стандарт MPI.
+
+Сравнение алгоритмов глобальной оптимизации проведено путём оценки качества решения алгоритмами выборки задач из некоторого тестового класса.
+В данной статье рассматривается тестовый класс, порождаемый генератором GKLS \cite{Gaviano2003}.
+Данные генератор позволяет конструировать сложные многоэкстремальные задачи различной размерности.
+В работе рассматриваются выборки по 100 задач классов размерности 2, 3, 4, 5.
+Каждый класс имеет две степени сложности --- \textit{Simple} и \textit{Hard}.
+Параметры генератора для рассматриваемых классов приведены в \cite{Gaviano2003}.
+
+In order to evaluate the efficiency of an algorithm on a given set of 100 problems, we will use the
+operating characteristics \cite{grishaginClass}, which are defined by a set of
+points on the \((K, P)\) plane where \(K\) is the average number of search trials
+conducted before satisfying the termination condition when minimizing a function
+from a given class, and \(P\) is the proportion of problems solved successfully.
+If at a given \(K\), the operating characteristic of a method goes higher than one
+from another method, it means that at fixed search costs, the former method has a
+greater probability of finding the solution. If some value of \(P\) is fixed, and the
+characteristic of a method goes to the left from that of another method, the former
+method requires fewer resources to achieve the same reliability.
+
+\subsection{Сравнение последовательных разверток}
+С целью понять, обладает ли какой-либо из перечисленных ранее типов развёрток существенным преимуществом над другими, были построены операционные характеристики индексного метода с различными типами развёрток на классах GKLS 2d Simple и GKLS 3d Simple.
+The global minimum was considered to be found if the algorithm generates a trial point $y^k$ in the $\delta$-vicinity of the global minimizer, i.e.
+$\left\|y^k-y^\ast\right\|_\infty\leq\delta$.
+The size of the vicinity was selected as $\delta = 0.01\left\|b-a\right\|$.
+In case of GKLS $\delta=0.01$.
+
+Во всех экспериментах параметр плотности построения развёрток $m=12$.
+Минимальное значение пераметра надёжности \(r\) было найдено для каждого типа развёртки перебором по равномерной сетке с шагом \(0.1\).
+
+На классе GKLS 2d Simple при минимальном \(r\) неинъективная и гладкая развёртка обеспечивают более быструю сходимость (рис.
+\ref{fig:gkls2d_opt}).
+То же самое, наблюдается и при \(r=5.0\) (рис.
+\ref{fig:gkls2d_acc}).
+В последнем случае сдвиговая и вращаемая развёртки начинают отставать от остальных, т.к.
+значение \(r=5.0\) является завышенным для них.
+\begin{figure}[ht]
+    \centering
+    \subfloat[$r=5.0$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS2d_same_r_opt_pt_op.pdf}}\label{fig:gkls2d_acc}}
+    \subfloat[Minimal $r$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS2d_opt_pt_op.pdf}}\label{fig:gkls2d_opt}}
+    \caption{Operating characteristics on GKLS 2d Simple class}
+\end{figure}
+
+На классе GKLS 2d Simple при минимальном \(r\) неинъективная и множественные развёртки имеют значительное преимущество над единственной развёрткой.
+Значение \(r=4.5\) является завышенным для вращаемой и сдвиговой развёрток.
+
+\begin{figure}[ht]
+    \centering
+    \subfloat[$r=4.5$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS3d_same_r_opt_pt_op.pdf}}\label{fig:gkls3d_acc}}
+    \subfloat[Minimal $r$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS3d_opt_pt_op.pdf}}\label{fig:gkls3d_opt}}
+    \caption{Operating characteristics on GKLS 3d Simple class}
+\end{figure}
+
+\paragraph{Накладные расходы при использовании сдвиговой развёртки.}
+
 \subsection{Параллельные вращаемые развертки}
-s
-\subsection{Параллельные вращаемые развертки для задач большой размерности}
-s
+
 \section{Вывод о целесообразности применения того или иного вида разверток для того или иного вида задач}
@@
 Strongin, R.G., Gergel, V.P., Barkalov, K.A.: Parallel methods for global optimization problem solving. Journal of instrument engineering. \textbf{52}, 25--33 (2009) (In Russian)
+
+\bibitem{Gaviano2003}
+Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
+ACM Transactions on Mathematical Software 29(4), 469--480 (2003)
+
+\bibitem{grishaginClass}
+Grishagin, V.A.: Operating Characteristics of Some Global Search Algorithms. Problems of Statistical Optimization 7, 198--206 (1978) (In Russian)
+
 \end{thebibliography}

### 2018-04-07 2018/RSD/MultiEvolvents/RSD.tex
RSD: add draft of tables in parallel section

@@
 $\left\|y^k-y^\ast\right\|_\infty\leq\delta$.
-The size of the vicinity was selected as $\delta = 0.01\left\|b-a\right\|$.
+The size of the vicinity was selected as $\delta = 0.01\left\|b-a\right\|_\infty$.
 In case of GKLS $\delta=0.01$.
@@
 
-На классе GKLS 2d Simple при минимальном \(r\) неинъективная и множественные развёртки имеют значительное преимущество над единственной развёрткой.
-Значение \(r=4.5\) является завышенным для вращаемой и сдвиговой развёрток.
+На классе GKLS 2d Simple при минимальном \(r\) неинъективная и множественные развёртки имеют значительное преимущество над единственной развёрткой (рис.
+\ref{fig:gkls3d_opt}).
+Значение \(r=4.5\) является завышенным для вращаемой и сдвиговой развёрток (рис.
+\ref{fig:gkls3d_acc}).
 
@@
 \paragraph{Накладные расходы при использовании сдвиговой развёртки.}
+Во всех представленных выше экспериментах при построении операционной характеристики учитывалось количество вычислений целевой функции из класса GKLS, однако в случае сдвиговой развёртки индексный метод решает задачу с ограничением \(g_0\) из (\ref{6_g0}).
+В точках, где \(g_0\) нарушено, значение целевой функции не вычисляется.
+Эти точки, тем не менее, хранятся в поисковой информации, создавая дополнительные расходы вычислительных ресурсов.
+В таблице \ref{tab:shifted_g0} приведено среднее количество обращение к \(g_0\) и целевой функции.
+При \(L=3\) ограниение \(g_0\) вычисляется почти в 20 раз чаще, чем целевая функция \(\varphi\), т.
+е.
+\(95\%\) всей поисковой информации приходится на вспомогательные точки.
+Такие накладные расходы приемлемы при решении задач малой размерности с трудоёмкими целевыми функциями, но при росте размерности и общего количества испытаний выгоднее использовать другие типы развёрток.
+
+\begin{table}
+\begin{center}
+\caption{Среднее количество вычислений \(g_0\) и \(\varphi\) при решении задач класса GKLS 3d Simple с помощью сдвиговой развёртки}
+  \begin{tabular}{|l|{c}|{c}|{c}|}
+    \hline
+  $L$ & Constraint & Objective (shifted) & $\frac{calc(g_0)}{calc(\varphi)}$ ratio \\
+  \hline
+  2 & 96247.9  & 6840.14 & 14.07\\
+  \hline
+  3 & 153131.0 & 7702.82 & 19.88\\
+  \hline
+  \end{tabular}
+  \label{tab:shifted_g0}
+\end{center}
+\end{table}
 
 \subsection{Параллельные вращаемые развертки}
+\begin{table}
+  \centering
+  \caption{Averaged numbers of iterations executed by GPMAGS for solving the test optimization problems}
+  \label{tab:4}
+  \begin{tabular}{cccccccc}
+    \cline{3-8}\noalign{\smallskip}
+    \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}   \\
+    \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
+    \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
+    \noalign{\smallskip}\hline
+    I &
+    \parbox{0.25\textwidth}{
+    \begin{center}
+    \textbf{1 cluster node}
+    \end{center}		}
+      & 1 & 11953 & 25263 & & 15920 & \(>\)148342 (4)  \\
+    &  & 32 & 11953 & 25263 & & 15920 & \(>\)148342 (4)  \\
+    \hline \noalign{\smallskip}
+II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
+  & 1 & 4762 & 11178 & & 13378 & 109075 \\
+&   & 32 & 2372 & 5972 & & 5203 & 51868 \\
+    \noalign{\smallskip}\hline	\noalign{\smallskip}
+III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
+  & 1  & 171 & 393 & & 382 & 3452  \\
+&    & 32 & 85 & 182 & & 249 & 1306 \\
+
+    \noalign{\smallskip}\hline
+  \end{tabular}
+\end{table}
+
+\begin{table}
+  \centering
+  \caption{Speedup of parallel computations executed by GMAGS}
+  \label{tab:4}
+  \begin{tabular}{cccccccc}
+    \cline{3-8}\noalign{\smallskip}
+    \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}   \\
+    \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
+    \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
+    \noalign{\smallskip}\hline
+    I &
+    \parbox{0.25\textwidth}{
+    \begin{center}
+    \textbf{1 cluster node}
+    \end{center}		}
+      & 1 & 11953 & 25263 & & 15920 & \(>\)148342 (4)  \\
+    &  & 32 & 11953 & 25263 & & 15920 & \(>\)148342 (4)  \\
+    \hline \noalign{\smallskip}
+II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
+  & 1 & 4762 & 11178 & & 13378 & 109075 \\
+&   & 32 & 2372 & 5972 & & 5203 & 51868 \\
+    \noalign{\smallskip}\hline	\noalign{\smallskip}
+III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
+  & 1  & 171 & 393 & & 382 & 3452  \\
+&    & 32 & 85 & 182 & & 249 & 1306 \\
+
+    \noalign{\smallskip}\hline
+  \end{tabular}
+\end{table}
 

### 2018-04-08 2018/RSD/MultiEvolvents/RSD.tex
RSD: update numerical experiments

@@
 In this paper, the core class of optimization problems, which can be solved using
-Globalizer, is formulated. This class involves the multidimensional global
+Globalizer\cite{globalizerSystem}, is formulated. This class involves the multidimensional global
 optimization problems without constraints, which can be defined in the following way:
@@
 Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. The CPUs had 8 cores (i. e. total 16 cores
-were available per a node). Все рассматриваемые алгортимы и развёртки были реализованы на языке C++ в рамках программной системы Globalizer.
+were available per a node). Все рассматриваемые алгортимы и развёртки были реализованы на языке C++ в рамках программной системы Globalizer\cite{globalizerSystem}.
 Для реализации параллелизма на одном узле использована технология OpenMP, а для параллелизма на несколько узлов --- стандарт MPI.
@@
     \hline
-  $L$ & Constraint & Objective (shifted) & $\frac{calc(g_0)}{calc(\varphi)}$ ratio \\
+  $L$ & $calc(g_0)$ & $calc(\varphi)$ & $\frac{calc(g_0)}{calc(\varphi)}$ ratio \\
   \hline
@@
 \subsection{Параллельные вращаемые развертки}
+Чтобы оценить эффективность параллельного алгортима из секции \ref{sec:parallel_evolvents} были проведены вычислительные эксперименты на классах GKLS 4d (Hard, Simple) и GKLS 5d (Hard, Simple).
+Значения \(r\) во всех экспериментах равно 5.0, размер \(\delta\)-окрестности известного решения увеличен до 0.3.
+При решении серий задач использовалось до 8 узлов кластера и до 32 вычислительных потоков на каждом узле.
+
+В таблице \ref{tab:iterations} приведено среднее количество итераций при решении 100 задач из каждого из рассматриваемых классов.
+При увеличении числа узлов и числа потоков на каждом узле количество итераций заметно сокращается (за исключением класса GKLS 4d Simple при переходе с 1 узла на 4 узла в однопоточном режиме).
+
 \begin{table}
   \centering
-  \caption{Averaged numbers of iterations executed by GPMAGS for solving the test optimization problems}
-  \label{tab:4}
+  \caption{Averaged numbers of iterations executed by the parallel algorithm for solving the test optimization problems}
+  \label{tab:iterations}
   \begin{tabular}{cccccccc}
@@
     \end{center}		}
-      & 1 & 11953 & 25263 & & 15920 & \(>\)148342 (4)  \\
-    &  & 32 & 11953 & 25263 & & 15920 & \(>\)148342 (4)  \\
+      & \textit{1} & 12167 & 25635 & & 20979 & 187353  \\
+    &  & \textit{32} & 328 & 1268  & &   898 & 12208 \\
     \hline \noalign{\smallskip}
 II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
-  & 1 & 4762 & 11178 & & 13378 & 109075 \\
-&   & 32 & 2372 & 5972 & & 5203 & 51868 \\
+  & \textit{1} & 25312 & 11103 & & 1472 & 17009 \\
+&   & \textit{32} & 64 &   913 & & 47 & 345 \\
     \noalign{\smallskip}\hline	\noalign{\smallskip}
 III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-  & 1  & 171 & 393 & & 382 & 3452  \\
-&    & 32 & 85 & 182 & & 249 & 1306 \\
-
+  & \textit{1}  & 810 & 4351 & & 868 & 5697  \\
+& & \textit{32} & 34  & 112  & & 35  & 868 \\
     \noalign{\smallskip}\hline
@@
 
+Если считать, что затраты на параллелизм пренебрежимо малы по сравнению с затратами на вычисление целевых функций в задачах оптимизации, то ускорение по времени от использования параллельного метода будет равно ускорению по итерациям.
+Однако, в действительности это предположение не всегда справедливо.
+Во всех численных экспериментах время вычисления целевой функции занимает примерно $10^{-3}$с.
+В таблице \ref{tab:speedup} приведено ускорение по итерациям и в куглых скобках ускорение по времени.
+В первой строке таблицы, соответствующей последовательному режиму, в скобках приведено среднее время решения одной задачи.
+Из таблицы видно, что для класов GKLS 4d выгоднее использовать один узел в многопоточном режиме, тогда как для решения более сложных пятимерных задач лучше использовать несколько узлов, каждый из которых работает в параллельном режиме.
+
 \begin{table}
   \centering
-  \caption{Speedup of parallel computations executed by GMAGS}
-  \label{tab:4}
+  \caption{Speedup of parallel computations executed by the parallel algorithm}
+  \label{tab:speedup}
   \begin{tabular}{cccccccc}
@@
     \end{center}		}
-      & 1 & 11953 & 25263 & & 15920 & \(>\)148342 (4)  \\
-    &  & 32 & 11953 & 25263 & & 15920 & \(>\)148342 (4)  \\
+    & \textit{1}   & 12167(10.58s) & 25635(22.26s) & & 20979(22.78s) & 187353(205.83s)  \\
+  &  & \textit{32} & 37.1(18.03) & 20.2(8.55)  & &  23.3(8.77) & 15.4(9.68) \\
   \hline \noalign{\smallskip}
 II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
-  & 1 & 4762 & 11178 & & 13378 & 109075 \\
-&   & 32 & 2372 & 5972 & & 5203 & 51868 \\
+& \textit{1} &        0.5(0.33) & 2.3(0.86)  & & 14.3(6.61) & 11.0(6.06) \\
+&   & \textit{32} & 190.1(9.59) & 28.1(1.08) & & 446.4(19.79) & 543.0(43.60) \\
   \noalign{\smallskip}\hline	\noalign{\smallskip}
 III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-  & 1  & 171 & 393 & & 382 & 3452  \\
-&    & 32 & 85 & 182 & & 249 & 1306 \\
-
+& \textit{1}    & 15.0(6.05)  & 5.9(2.36)   & & 24.2(17.56)  & 32.9(24.87)  \\
+& & \textit{32} & 357.9(2.36) & 228.9(2.64) & & 582.8(20.96) & 793.0(33.89) \\
     \noalign{\smallskip}\hline
@@
 \section{Вывод о целесообразности применения того или иного вида разверток для того или иного вида задач}
-s
 
@@
 
+\bibitem{globalizerSystem}
+Gergel V.P., Barkalov K.A., and Sysoyev A.V: Globalizer: A novel supercomputer software system for solving time-consuming global optimization problems.
+Numerical Algebra, Control \& Optimization 8(1), 47--62, 2018
+
 \end{thebibliography}

### 2018-04-08 2018/RSD/MultiEvolvents/RSD.tex
RSD: add conclusion

@@
 \end{abstract}
-%
-%
-%
 
+%------------------------------------------------------------------------------
 \section{Introduction}\label{sec:intro}
@@
 
+%------------------------------------------------------------------------------
 \section{Statement of Multidimensional Global Optimization Problem}
@@
 
+%------------------------------------------------------------------------------
 \section{Methods of Dimension Reduction}
@@
 
+%------------------------------------------------------------------------------
 \subsection{Сдвиговые}
@@
 
+%------------------------------------------------------------------------------
 \subsection{Вращаемые}
@@
 
+%------------------------------------------------------------------------------
 \subsection{Non injective evolvent}
@@
 
+%------------------------------------------------------------------------------
 \subsection{Smooth evolvent}
@@
 
+%------------------------------------------------------------------------------
 \subsection{Параллельные множественные отображения}
@@
 
+%------------------------------------------------------------------------------
 \section{Results of Numerical Experiments}
@@
 
+%------------------------------------------------------------------------------
 \subsection{Сравнение последовательных разверток}
+\label{sec:seq_comp}
 С целью понять, обладает ли какой-либо из перечисленных ранее типов развёрток существенным преимуществом над другими, были построены операционные характеристики индексного метода с различными типами развёрток на классах GKLS 2d Simple и GKLS 3d Simple.
@@
 
+%------------------------------------------------------------------------------
 \subsection{Параллельные вращаемые развертки}
+\label{sec:results_parallel}
 Чтобы оценить эффективность параллельного алгортима из секции \ref{sec:parallel_evolvents} были проведены вычислительные эксперименты на классах GKLS 4d (Hard, Simple) и GKLS 5d (Hard, Simple).
@@
 
-\section{Вывод о целесообразности применения того или иного вида разверток для того или иного вида задач}
+%------------------------------------------------------------------------------
+\section{Conclusions}
+В работе рассмотрены 5 различных отображений типа кривой Пеано, применяемых для редукции размерности в задачах глобальной оптимизации.
+Из предварительного сравнения, проведенного в секции \ref{sec:seq_comp}, можно сделать слудующие выводы:
+\begin{itemize}
+  \item гладкая и неинъективная развёртки показывают наилучший результат в задачах малой размерности и могут быть успешно применены при решении задач с трудоемкими целевыми функциями.
+Особенности этих развёрток не позволяют создать на их основе масштабируемый на несколько узлов кластера алгоритм оптимизации.
+  \item сдвиговые развёртки привносят большие накладные расходы на работу метода из-за того, что требуется добавления в задачу (\ref{eq:task}) вспомогательного функционального ограничения.
+Эксперименты показали, что до 95\% поисковой информации приходится на точки, в которых вичисляется только вспомогательное ограничение.
+Сдвиговые развёртки могут быть использованы в качестве основы для параллельного алгоритма из секции \ref{sec:parallel_evolvents}, однако расходы на обработку дополнительных точек, скорее всего, приведут к малому коэффициенту ускорения от параллельности.
+Однако, при достаточно трудоёмкой целевой функции их использование может иметь смысл.
+  \item вращаемые развёртки обеспечили приемлемую скорость сходимости на задачах малой размерности в последовательном случае.
+Их использование не приводит к введению дополнительных ограничений, что позволяет построить эффективный параллельный алгоритм на их основе.
+\end{itemize}
+
+В секции \ref{sec:results_parallel} приведены результаты численных экспериментов, которые показывают, что алгоритм из раздела \ref{sec:parallel_evolvents} на основе вращаемых развёрток позволяет получить ускорение до 43 раз при решении серии задач с использованием нескольких узлов вычислитеьного кластера.
+Стоит заметить, что в расматриваемых задачах целевые функции не являются трудоёмкими (среднее время вычисления $10^{-3}$с) и в случае более сложных задач показатели ускорения по времени будут приближаться к ускорению по итерациям.
 

### 2018-04-08 2018/RSD/MultiEvolvents/RSD.tex
RSD: tiny fixes

@@
 This allows problems to be solved in which function values may not be determined for the entire search domain.
-Under this approach, solving multidimen-sional problems is reduced (using Peano-type space-filling curves) to solving equivalent one-dimensional problems.
+Under this approach, solving multidimensional problems is reduced (using Peano-type spacefilling curves) to solving equivalent one-dimensional problems.
 It should be noted that standard approaches to algorithm parallelization are not quite applicable to global optimization.
-For example, the rules for selecting an-other iteration point are quite simple and do not require parallelization (as over-heads associated with organizing parallel computations will nullify any possible acceleration).
-Some acceleration can be achieved by parallelizing the computa-tion of function values describing the object to be optimized; however, this ap-proach is specific to each individual problem being solved.
+For example, the rules for selecting another iteration point are quite simple and do not require parallelization (as overheads associated with organizing parallel computations will nullify any possible acceleration).
+Some acceleration can be achieved by parallelizing the computa-tion of function values describing the object to be optimized; however, this approach is specific to each individual problem being solved.
 The following approach looks more promising.
 The algorithm can be modified to run several trials in parallel.
-This approach provides the efficiency (as paral-lelization is applied to the most computation-intensive part of the problem solv-ing process) and generality (in that it applies to a wide range of global optimiza-tion algorithms).
+This approach provides the efficiency (as parallelization is applied to the most computation-intensive part of the problem solving process) and generality (in that it applies to a wide range of global optimiza-tion algorithms).
 The approach, described in  \cite{Two_Level_Parallel} for unconstrained optimization, was used in this work for parallelizing constrained optimization algorithms.
@@
 \section{Methods of Dimension Reduction}
-\subsection{Базовый алгоритм}
+\subsection{Single evolvent}
 
@@
 %------------------------------------------------------------------------------
-\subsection{Сдвиговые}
+\subsection{Shifted evolvents}
 \label{sec:shifted}
@@
 %------------------------------------------------------------------------------
-\subsection{Вращаемые}
+\subsection{Rotated evolvents}
 The application of the scheme for building the multiple evolvents (hereinafter called the shifted evolvents or $S$-evolvents) described in Subsection \ref{sec:shifted} allows to preserve the information on the nearness of the points in the multidimensional space and, therefore, to provide more precise (as compared to a single evolvent) estimate of Lipschitz constant in the search process.

### 2018-04-08 2018/RSD/MultiEvolvents/RSD.tex
RSD: drop the old introduction. We need a more suitable one

@@
 % \renewcommand\UrlFont{\color{blue}\rmfamily}
+\setlength{\textfloatsep}{4pt plus 1.0pt minus 1.5pt}
+\setlength{\floatsep}{4pt plus 1.0pt minus 1.5pt}
+%\setlength{\abovedisplayskip}{3pt}
+%\setlength{\belowdisplayskip}{3pt}
 
@@
 
-Global (or multiextremal) optimization problems are among the most complex problems
-in both theory and practice of optimal decision making. In these kinds of problems,
-the optimization criterion can have several local optima within the search domain.
-The existence of several local optima makes
-finding the global optimum difficult essentially, since it requires examining the whole feasible
-search domain. The volume of computations for solving global optimization problems
-can increase exponentially with increasing number of varied parameters.
-\par
-These global optimization problem features impose special requirements on the quality
-of the optimization methods and on the software to implement these ones. The global
-optimization methods should be highly efficient, and the software systems should be developed on
-a good professional basis. In general, the global optimization problems can be solved at
-a reasonable time by employing parallel computations on modern supercomputing systems only.
-\par
-The general state of the art in the field of global optimization is presented in a
-number of key monographs  \cite{floudasPardGO}, \cite{horstTuyGO}, \cite{locatelliSchoenGO}, \cite{pinterGO}, \cite{strSergGO}, \cite{zilinskTornGO}, \cite{zhigljavskyRandGO}.
-The development of optimization methods, which use the high-performance computational
-systems to solve the time-consuming global optimization problems, is an area of
-intensive research --- see, for instance, \cite{censorZeniosParGO}, \cite{ciegisHentyParGO},
-\cite{luqueAlbaGA}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}. The obtained
-theoretical results provide the efficient solutions of many applied global
-optimization problems in various fields of scientific and technical applications \cite{Famularo1999}, \cite{fasanoPinter2013},
-\cite{floudasPardalosGOState}, \cite{Kvasov2015}, \cite{Menniti}, \cite{locatelliSchoenGO},
-\cite{luqueAlbaGA}, \cite{pardalosZhigljavskyZilinskas2016}, \cite{pinterGO}.
-\par
-At the same time, the practical implementation of these global optimization algorithms
-within the framework of industrial software systems is quite limited. In many cases,
-software implementations are experimental in nature and are used by the developers
-themselves to obtain the results from the computational experiments required for the
-scientific publications. This situation originates from high development costs of the
-professional software systems, which can be used by numerous users. In addition, the global
-optimization problems could be solved in an automatic mode rarely because of the
-complexity of these ones. The user should actively control the global search
-process that implies an adequate level of qualification in the field of optimization
-(particularly, the user should know and understand the global optimization methods well).
-\par
-In this work, the authors consider an approach to minimizing multiextremal functions developed in ...????????
-This allows problems to be solved in which function values may not be determined for the entire search domain.
-Under this approach, solving multidimensional problems is reduced (using Peano-type spacefilling curves) to solving equivalent one-dimensional problems.
-It should be noted that standard approaches to algorithm parallelization are not quite applicable to global optimization.
-For example, the rules for selecting another iteration point are quite simple and do not require parallelization (as overheads associated with organizing parallel computations will nullify any possible acceleration).
-Some acceleration can be achieved by parallelizing the computa-tion of function values describing the object to be optimized; however, this approach is specific to each individual problem being solved.
-The following approach looks more promising.
-The algorithm can be modified to run several trials in parallel.
-This approach provides the efficiency (as parallelization is applied to the most computation-intensive part of the problem solving process) and generality (in that it applies to a wide range of global optimiza-tion algorithms).
-The approach, described in  \cite{Two_Level_Parallel} for unconstrained optimization, was used in this work for parallelizing constrained optimization algorithms.
-
 %------------------------------------------------------------------------------
@@
   \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
-  D=\{y\in \mathbf{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
+  D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
 \end{array}
@@
 basis of a finite number of computed values of the optimized function only in this case .
-\par
-As it has been shown earlier by many researchers
-(see, for instance, \cite{floudasPardalosGOState}, \cite{horstTuyGO}, \cite{pinterGO}, \cite{strSergGO}),
-finding the numerical estimate of the global optimum implies constructing a coverage of
-the search domain \(D\). As a result, the computational costs of solving the global
-optimization problems are readily very high even for a small number of varied parameters
-(the dimensionality of the problem). A notable reduction in the volume of computations
-can be achieved when the coverage of the search domain is non-uniform, i. e. the series
-of trial points is only dense in a vicinity of the global optimum point. The construction
-of such a non-uniform coverage could be provided in an adaptive way, when the selection
-of the next trial points is determined by using the search information (the preceding
-trial points and the values of the minimized function at these points) obtained in the course
-of computations. This necessary condition complicates considerably the computational schemes
-of global optimization methods since it implies a complex analysis of a large amount
-of multidimensional search information. As a result, many optimization algorithms use
-various approaches to the dimensional reduction \cite{pinterGO}, \cite{sergeyevStronginLera2013}, \cite{strongin1978}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
 
@@
 The algorithms for the numerical construction of the Peano curve approximations are
-given in \cite{strSergGO}. As an illustration, an approximation of the Peano curve
-for the third density level is shown in Figure \ref{fig:peanoC}. Figure \ref{fig:peanoC}
-demonstrates the movement order in a two-dimensional domain to construct the Peano
-curve approximation; the precision of the Peano curve approximation is determined by the
-density level used in the construction.
-\begin{figure}
-    \centering
-    \includegraphics[width=0.85\textwidth]{pictures/peanoC.eps}
-    \caption{A Peano curve approximation for the third density level}
-    \label{fig:peanoC}
-\end{figure}
+given in \cite{strSergGO}.
 
@@
 The computational scheme obtained as a result of the dimensionality reduction consists of the following:
-%(see Figure \ref{fig:peanoCUsage}):
 \begin{itemize}
@@
 \end{itemize}
-
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=0.55\textwidth]{pictures/peanoCUsage.eps}
-%    \caption{The computational scheme for obtaining the value of the reduced one-dimensional function \(\varphi(y(x))\)}
-%    \label{fig:peanoCUsage}
-%\end{figure}
 
@@
 
-Let us consider a family of the hypercubes
-\begin{equation}\label{6_hypercubes}
-D_l= \left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq L,
-\end{equation}
-where the hypercube $D_{l+1}$ is obtained by the translation of the hypercube $D_l$ along the main diagonal with displacement equal to $2^{-l}$ in each coordinate.
-In Fig.~\ref{fig:shifted_ev} the hypercubes $D_0,...,D_3$ for the case $N=2,\ L=3$ are presented.
+Such set of evolvents can be produced by shifting the source evolvent $y^0(x)$ by $2^{-l},0 \leq l \leq L$ on each coordinate.
+Each evolvent has it's own corresponding hypercude $D_l= \left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq L$.
 
-\begin{figure}[ht]
-    \centering
-    \subfloat[Two mappings]{{\includegraphics[width=.6\textwidth]{pictures/shifted.pdf}}\label{fig:shifted_ev}}
-    \subfloat[Hypercubes $D_l$]{{\includegraphics[width=.4\textwidth]{pictures/shifted_cube.png}}\label{fig:shifted_cube}}
-    \caption{Multiple mappings}
-\end{figure}
-
-Let us assume that the evolvent $y^0(x)$  maps the interval $[0,1]$ onto the hypercube $D_0$ from (\ref{6_hypercubes}), i.e.,
-\[
-D_0 = \left\{y^0(x) : x \in [0,1]\right\}.
-\]
-Then, the evolvents $y^l(x)=\left\{y_1^l(x),...,y_N^l(x)\right\}$, the coordinates of which are defined by the conditions
-\[
-y_i^l(x)=y_i^{l-1}(x)+2^{-l},\ 1\leq i\leq N, \ 1\leq l\leq L,
-\]
-map the interval $[0,1]$ onto the corresponding hypercubes $D_l,\ 1\leq l \leq L$.
 In Fig.~\ref{fig:shifted_ev} the image of the interval $[0,1]$ obtained by the curve $y^0(x),\ x\in [0,1],$ is shown as the dashed line.
-Since the hypercube $D$ from (\ref{eq:task}) is included in the common part of the family of hypercubes (\ref{6_hypercubes}) (the boundaries of hypercube $D$ are highlighted in Fig.~\ref{fig:shifted_cube}), having introduced an additional constraint function
+Since the hypercube $D$ from (\ref{eq:task}) is included in the common part of the family of hypercubes $D_l$ (the boundaries of hypercube $D$ are highlighted in Fig.~\ref{fig:shifted_cube}), having introduced an additional constraint function
 \begin{equation}\label{6_g0}
@@
 
-The problems (\ref{6_problem_l}) correspond to the domains $Q_0^l=[0,1]$ and $Q_{j+1}^l, 0 \leq j \leq m,$ defined by the expression
-\[
-Q_{j+1}^l = \left\{x \in Q_j^l:g_j(y^l(x))\leq 0\right\},\ 0\leq j\leq m,
-\]
-for the corresponding evolvents $y^l(x)$. The application of a multiple mapping defines the following relation for the nearness in the multidimensional search domain and in the one-dimensional one.
-
-\begin{theorem}
-Let a point $y^\ast$ from the domain $D$ be contained in the line segment with the end-points $y',y'' \in D$ meeting the requirements
-\[
-\left|y'_j - y''_j\right|\leq 2^{-p},\ y'_i=y''_i=y^\ast_i, \ 1\leq i \leq N,\ i \neq j,
-\]
-where $p$ is an integer  and $1\leq p \leq L$, i.e., the line segment is collinear with the $j$-th axis in $R^N$.
-Then, there exist at least one mapping $y^l(x),\ 0\leq l\leq L$, and the preimages $x^\ast,\; x',\; x''\in [0,1]$ such that
-\[
-y^\ast = y^l(x^\ast),\ y'=y^l(x'),\ y'' = y^l(x'')
-\]
-and
-\[
-\max \left\{ \left|x'-x^\ast\right|,\; \left|x''-x^\ast\right|,\; \left|x'-x''\right|\right\} \leq 2^{-pN}.
-\]
-\end{theorem}
-\begin{proof}
-Proof of this theorem is given in \cite{strSergGO}.
-\qed
-\end{proof}
-
-The conditions of the theorem single out a specific vicinity of the point $y^\ast$.
-This vicinity includes only the points, which can be obtained by the shift of $y^\ast$ parallel to one of the coordinate axes with a displacement not more than $2^{-p}$.
-By changing  $j,\ 1\leq j\leq N,$ in the theorem conditions it is possible to obtain the neighbours in any $N$ coordinate directions.
-According to the statement, the closeness of the points in the $N$-dimensional space in a particular direction will be reflected by the closeness of their preimages in one of the univariate problems.
-The information on the closeness of the points results, first, in more precise estimate of Lipschitz constants and, second, in the increase of the characteristics of the intervals, the images of the end points of which are close to each other in the $N$-dimensional space.
+\begin{figure}[ht]
+    \centering
+    \subfloat[Two mappings]{{\includegraphics[width=.6\textwidth]{pictures/shifted.pdf}}\label{fig:shifted_ev}}
+    \subfloat[Hypercubes $D_l$]{{\includegraphics[width=.4\textwidth]{pictures/shifted_cube.png}}\label{fig:shifted_cube}}
+    \caption{Multiple mappings}
+\end{figure}
 
@@
 The application of the scheme for building the multiple evolvents (hereinafter called the shifted evolvents or $S$-evolvents) described in Subsection \ref{sec:shifted} allows to preserve the information on the nearness of the points in the multidimensional space and, therefore, to provide more precise (as compared to a single evolvent) estimate of Lipschitz constant in the search process.
-However, this approach has serious restrictions, which narrow the applicability of the parallel algorithms, designed on the base of the $S$-evolvents.
+However, this approach has serious restrictions, which narrow the applicability of the parallel algorithms, designed on the base of the $S$-evolvents (see te end of the section \ref{sec:seq_comp}).
 
-Because a shifted evolvent is built as a result of the shift of a hypercube $D$ along the main diagonal, and the shift step decreases 2 times for the building of each subsequent mapping, the number of such shifts is limited by the evolvent density $m$.
-In the case when the shift step is less than $2^{-m}$, the next evolvent coincides with the previous one.
-Thus, the applicable number of the evolvents $L$ and, hence, the number of processors are limited by $m$ ($L\leq m$), where $m$ is the density of the evolvents.
-
-Also, it follows from the algorithm for building the shifted evolvents that solving the initial problem in the domain $D$ is reduced to solving a set of problems in the extended domains $D_l$ from (\ref{6_hypercubes}) with additional constraint (\ref{6_g0}).
-As a result, the structure of the search domain becomes more complicated, and the exponential decrease of the volume of the search domain $D$ in relation to the volumes of the domains $D_l$ with increasing dimensionality of the problem takes place.
-
-To overcome the disadvantages mentioned above and to preserve the information on the nearness of the points in the $N$-dimensional space, a novel scheme of building of the multiple mappings is proposed.
+To overcome the $S$-evolvent and to preserve the information on the nearness of the points in the $N$-dimensional space, a novel scheme of building of the multiple mappings is proposed.
 The building of a set of Peano curves not by the shift along the main diagonal of the hypercube but by rotation of the evolvents around the coordinate origin is a distinctive feature of the proposed scheme \cite{Gergel2009}.
@@
 In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer.
-The evolvents generated according to the novel scheme will be hereinafter called the rotated evolvents or $R$-evolvents.
+The evolvents generated according to this scheme are called the rotated evolvents or $R$-evolvents.
 
 In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case $N=2$ are presented as an illustration.
+Taking into account the initial mapping, one can conclude that current implementation of the method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals.
+Moreover, the additional constraint  $g_0(y) \leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent.
+This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.
 
@@
 \end{figure}
-
-Taking into account the initial mapping, one can conclude that current implementation of the method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals.
-Moreover, the additional constraint  $g_0(y) \leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent.
-This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.
-
-The use of the set of mappings $Y_L(x)=\{y^1(x),...,y^L(x)\}$ results in the appearing of the corresponding set of the one-dimensional multiextremal problems
-\begin{equation}\label{6_problem_lr}
-  \min{\left\{\varphi(y^l(x)):x\in [0,1], \; g_j(y^l(x))\leq 0, \; 1 \leq j \leq m\right\}}, \ 1 \leq l \leq L.
-\end{equation}
-Each problem from this set can be solved independently.
-Any computation of the value $z=g_\nu(y'),\ y'=y^i(x')$ of the function $g_\nu(y)$ in the $i$-th problem can be interpreted as a computation of the value $z=g_\nu(y'),\ y'=y^s(x'')$ for any other $s$-th problem without time-comsuming computations of the functions $g_\nu(y)$.
-Such information integrity allows to solve  initial problem (\ref{eq:task}) by solving $L$ problems (\ref{6_problem_lr}) on a set of intervals $[0,1]$ in parallel by the index method.
 
@@
 
-%\begin{figure}[t]
-%  \centering
-%  \includegraphics[width=0.6\linewidth]{pictures/noninjective.pdf}
-%  \caption{Numerical non injective evolvent}
-%  \label{fig:noninjective}
-%\end{figure}
-
 Недостатком неинъективной развёртки является потенциально большое количетво прообразов (до $2^N$) и невозможность использования параллельной схемы для множественных отображений из секции \ref{sec:parallel_evolvents}.
@@
 \subsection{Smooth evolvent}
-
-%\begin{figure}[t]
-%  \centering
-%  \includegraphics[width=0.6\linewidth]{pictures/smooth.pdf}
-%  \caption{Smooth evolvent}
-%  \label{fig:smooth}
-%\end{figure}
 
@@
 
-The information-statistical theory of global optimization formulated in \cite{strongin1978}, \cite{strSergGO} has
-served as a basis for the development of a large number of efficient multiextremal optimization
-methods \cite{barkalovGergel2014}, \cite{gergel1996}, \cite{gergel1997}, \cite{grishaginStrongin1984}, \cite{Pizzuti}, \cite{sergeyev1995}, \cite{sergeyev1999}, \cite{sergeyevGrishagin2001}, \cite{sergeyevStronginLera2013}, \cite{Famularo2001}.
-\par
-The optimization methods applied in Globalizer are based on the MAGS method, which can be presented as follows --- see \cite{strongin1978}, \cite{strSergGO}.
-\par
-Let us introduce a simpler notation for the optimization problem being solved
-\begin{equation}
-\label{eq:oneDimFunc}
-f(x) = \varphi(y(x)):x\in [0,1].
-\end{equation}
+The optimization methods applied in Globalizer to solve the reduced problem (\ref{eq:oneDimTask}) are based on the MAGS method, which can be presented as follows --- see \cite{strongin1978}, \cite{strSergGO}.
 \par
@@
 The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the following rules.
-\par
+
 \textit{Rule 1}. Renumber the points of the preceding trials by the lower indices in order of increasing value of coordinates
-\begin{equation}
-  \label{step1}
-0=x_0<x_1<...<x_{k+1}=1,
-\end{equation}
-the points \(x_0\), \(x_{k+1}\) were introduced additionally for the convenience of further
-explanation, the values of the minimized function \(z_0\), \(z_{k+1}\) at these points are undefined.
-\par
-\textit{Rule 2}. Compute the current estimate of the H{\"o}lder constant \(H\) from (\ref{eq:holder})
-\begin{equation} \label{step2}
-M=\max_{1 < i\leq k}\frac{|z_i-z_{i-1}|}{\rho_i}, \;
-m = \left\{
-   \begin{array}{lr}
-     rM, & M > 0,\\
-     1, & M = 0,
-   \end{array}
-  \right.
-\end{equation}
-as the maximum of the relative differences of the minimized function values on the
-set of previously executed trial points \(x_i,1\leq i\leq k\) from (\ref{step1}).
-Hereafter, \(\rho_i=(x_i-x_{i-1})^\frac{1}{N},1\leq i\leq k+1\). The
-constant \(r\), \(r>1\), is the reliability parameter of the algorithm.
-\par
-\textit{Rule 3}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq k+1\), where
-\[
-R(i)=2\rho_i-4\frac{z_i}{m},\quad i=1,
-\]
-\begin{equation} \label{step3}
-R(i)=\rho_i+\frac{(z_i-z_{i-1})^2}{m^2\rho_i}-2\frac{z_i+z_{i-1}}{m},\quad 1<i<k+1, \\
-\end{equation}
-\[
-R(i)=2\rho_{i}-4\frac{z_{i-1}}{m},\quad i=k+1.
-\]
+$0=x_0<x_1<...<x_{k+1}=1$.
 
-\par
-\textit{Rule 4}. Determine the interval with the maximum characteristic
-\begin{equation} \label{step4}
-R(t)=\max_{1\leq i \leq k+1}R(i).
-\end{equation}
-\par
-\textit{Rule 5}. Execute a new trial at the point \(x^{k+1}\) located within the interval
-with the maximum characteristic from (\ref{step4})
-\begin{equation} \label{step5}
-  x^{k+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{r|z_{t}-z_{t-1}|}{m}\right]^N,\; \textrm{ if } 1<t<k+1,
-\end{equation}
-\[
-  x^{k+1}=\frac{x_t+x_{t-1}}{2},\; \textrm{ if } t=1 \textrm{ or } t=k+1.
-\]
+\textit{Rule 2}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq k+1\).
 
-\par
-The stopping condition, which terminated the trials, is defined by the inequality
-\begin{equation}
-  \label{eq:stop_1}
-\rho_t<\varepsilon
-\end{equation}
-for the interval with the maximum characteristic from (\ref{step4}) and \(\varepsilon >0\) is the predefined
+\textit{Rule 4}. Determine the interval with the maximum characteristic $R(t)=\max_{1\leq i \leq k+1}R(i)$.
+
+\textit{Rule 5}. Execute a new trial at the point \(x^{k+1}\) located within the interval with the maximum characteristic from the previous step
+  $x^{k+1}=d(x_t)$.
+
+The stopping condition, which terminated the trials, is defined by the inequality $\rho_t<\varepsilon$
+for the interval with the maximum characteristic from Step 4 and \(\varepsilon >0\) is the predefined
 accuracy of the optimization problem solution. If the stopping condition is not satisfied,
 the index \(k\) is incremented by 1, and the new global optimization iteration is executed.
-\par
-In order to explain the algorithm presented above, let us note the following.
-The characteristics \(R(i), 1\leq i\leq k+1\), calculated according to (\ref{step3}) could
-be interpreted as some measures of importance of the intervals with respect to the
-expected location of the global minimum point. Thus, the rules (\ref{step4}) and (\ref{step5}) for selecting
-the interval of the next trial become more clear --- the point of every next
-global optimization iteration is selected within the interval, where the global minimum
-point can be found most likely.
-\par
-The convergence conditions of the described algorithm are given, for example, in \cite{strSergGO}.
+
+The convergence conditions and exact formulas for descision rules $R(i)$ and $d(x)$ of the described algorithm are given, for example, in \cite{strSergGO}.
 
 %------------------------------------------------------------------------------
-\subsection{Параллельные множественные отображения}
+\subsection{Parallel algorithm exploiting a set of evolvents}
 \label{sec:parallel_evolvents}
-Using the multiple mapping allows solving initial problem (1) by parallel solving the problems
+Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the problems
 \[
-\min\{\varphi(y^s(x)):x\in [0,1]\}, 1≤s≤S
+\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
 \]
-on a set of intervals [0,1] by the index method.
+on a set of intervals $[0,1]$ by the index method.
 Each one-dimensional problem is solved on a separate processor.
-The trial results at the point \(x^k\) obtained for the problem being solved by particular processor are interpreted as the results of the trials in the rest problems (in the corresponding points \(x^(k_1 ),…,x^(k_S ))\).
+The trial results at the point \(x^k\) obtained for the problem being solved by particular processor are interpreted as the results of the trials in the rest problems (in the corresponding points \(x^(k_1 ),\dots,x^(k_S))\).
 In this approach, a trial at the point \(x^k \in [0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of operations.
@@
 
-\bibitem{barkalovGergel2014}
-\newblock K. Barkalov and V. Gergel,
-\newblock \emph{\emph{Multilevel scheme of dimensionality reduction for parallel global search algorithms}},
-\newblock in \emph{Proceedings of the 1st International Conference on Engineering and Applied Sciences Optimization}, (2014), 2111--2124.
-
-\bibitem{barkalovGergel2015}% (MR3539505) [10.1007/s10898-016-0411-y]
-\newblock K. Barkalov and V. Gergel,
-\newblock \emph{\emph{Parallel global optimization on GPU}},
-\newblock \emph{J. Glob. Optim.}, \textbf{66} (2016), 3--20.
-
-\bibitem{barkalovGergelLebedev2015} %[10.1007/978-3-319-21909-7_31]
-\newblock K. Barkalov, V. Gergel and I. Lebedev,
-\newblock \emph{\emph{Use of Xeon Phi coprocessor for solving global optimization problems}},
-\newblock \emph{LNCS}, \textbf{9251} (2015), 307--318.
-
-\bibitem{barkalovGergelLebedevSysoev2015}
-\newblock K. Barkalov, V. Gergel, I. Lebedev and A. Sysoev,
-\newblock \emph{\emph{Solving the global optimization problems on heterogeneous cluster systems}},
-\newblock in \emph{CEUR Workshop Proceedings}, \textbf{1482} (2015), 411--419.
-
-\bibitem{Barkalov2013}% [10.1007/978-3-642-39958-9_14]
-\newblock K. Barkalov, A. Polovinkin, I. Meyerov, S. Sidorov and N. Zolotykh,
-\newblock \emph{\emph{SVM regression parameters optimization using parallel global search algorithm}},
-\newblock \emph{LNCS}, \textbf{7979} (2013), 154--166.
-
-\bibitem{bussieckMeeraus2004} (MR3618583)% [10.1007/978-1-4613-0215-5_8]
-\newblock M. R. Bussieck and A. Meeraus,
-\newblock \emph{\emph{General algebraic modeling system (GAMS)}},
-\newblock in \emph{Modeling Languages in Mathematical Optimization} (ed. J. Kallrath), Springer, (2004), 137--157.
-
 \bibitem{censorZeniosParGO}% (MR1486040)
@@
 \newblock Springer, 2009.
-
-\bibitem{iosoDescription}
-\newblock I. N. Egorov, G. V. Kretinin, I. A. Leshchenko and S. V. Kuptzov,
-\newblock \emph{\emph{IOSO optimization toolkit --- novel software to create better design}},
-\newblock in \emph{9th AIAA/ISSMO Symposium on Multidisciplinary Analysis and Optimization}, 2002. Available from \url{http://www.iosotech.com/text/2002\_4329.pdf}.
 
@@
 
-\bibitem{gavianoKvasovLeraSergeev2003} %(MR2077342) [10.1145/962437.962444]
-\newblock M. Gaviano, D. E. Kvasov, D. Lera and Y. D. Sergeev,
-\newblock \emph{\emph{Software for generation of classes of test functions with known local and global minima for global optimization}},
-\newblock \emph{ACM Trans. Math. Software}, \textbf{29} (2003), 469--480.
-
-\bibitem{gergelLebedev2015}% [10.1016/j.procs.2015.11.008]
-\newblock V. Gergel and I. Lebedev,
-\newblock \emph{\emph{Heterogeneous parallel computations for solving global optimization problems}},
-\newblock \emph{Procedia Comput. Sci.}, \textbf{66} (2015), 53--62.
-
-\bibitem{gergel1993}
-\newblock V. Gergel,
-\newblock \emph{\emph{A software system for multi-extremal optimization}},
-\newblock \emph{Eur. J. Oper. Res.}, \textbf{65} (1993), 305--313.
-
-\bibitem{gergel1996}% (MR1399766)
-\newblock V. Gergel,
-\newblock \emph{\emph{A method for using derivatives in the minimization of multiextremum functions}},
-\newblock \emph{Comput. Math. Math. Phys.}, \textbf{36} (1996), 729--742.
-
-\bibitem{gergel1997} %(MR1443089) [10.1023/A:1008290629896]
-\newblock V. Gergel,
-\newblock \emph{\emph{A global optimization algorithm for multivariate functions with Lipschitzian first derivatives}},
-\newblock \emph{J. Glob. Optim.}, \textbf{10} (1997), 257--281.
-
-\bibitem{gergel2013} %[doi:10.1016/j.procs.2013.05.164]
-\newblock V. Gergel, et al.,
-\newblock \emph{\emph{High performance computing in biomedical applications}},
-\newblock \emph{Procedia Computer Science}, \textbf{18} (2013), 10--19.
-
-\bibitem{gergel2015}% [10.15866/ireaco.v8i1.4935]
-\newblock V. Gergel, et al.,
-\newblock \emph{\emph{Recognition of surface defects of cold-rolling sheets based on method of localities}},
-\newblock \emph{International Review of Automatic Control, } \textbf{8} (2015), 51--55.
-
-\bibitem{gergelSidorov2015}% [10.1007/978-3-319-21909-7_49]
-\newblock V. Gergel and S. Sidorov,
-\newblock \emph{\emph{A two-level parallel global search algorithm for solving computationally intensive multi-extremal optimization problems}},
-\newblock \emph{LNCS}, \textbf{9251} (2015), 505--515.
-
-\bibitem{grishaginStrongin1984}% (MR785497)
-\newblock V. A. Grishagin and R. G. Strongin,
-\newblock \emph{\emph{Optimization of multi-extremal functions subject to monotonically unimodal constraints}},
-\newblock \emph{Engineering Cybernetics}, \textbf{5} (1984), 117--122.
-
-\bibitem{holmstromEdvall2004}% [10.1007/978-1-4613-0215-5_19]
-\newblock K. Holmstrm and M. M. Edvall,
-\newblock \emph{\emph{The TOMLAB optimization environment}},
-\newblock \emph{Modeling Languages in Mathematical Optimization}, Springer, (2004), 369--376.
-
 \bibitem{horstTuyGO}% (MR1102239) [10.1007/978-3-662-03199-5]
@@
 \newblock Springer-Verlag, Berlin, 1990.
-
-\bibitem{Jones} (MR1246501)% [10.1007/BF00941892]
-\newblock D. R. Jones, C. D. Perttunen and B. E. Stuckman,
-\newblock \emph{\emph{Lipschitzian optimization without the Lipschitz constant}},
-\newblock \emph{J. Optim. Theory Appl.}, \textbf{79} (1993), 157--181.
-
-\bibitem{kearfott2009}%(MR2554906) [10.1080/10556780802614051]
-\newblock R. B. Kearfott,
-\newblock \emph{\emph{GlobSol user guide}},
-\newblock \emph{Optim. Methods Softw.}, \textbf{24} (2009), 687--708.
 
@@
 
-\bibitem{Pizzuti}% (MR1971214) [10.1007/s00211-002-0419-8]
-\newblock D. E. Kvasov, C. Pizzuti and Y. D. Sergeyev,
-\newblock \emph{\emph{Local tuning and partition strategies for diagonal GO methods}},
-\newblock \emph{Numerische Mathematik}, \textbf{94} (2003), 93--106.
-
-\bibitem{liberti2009}% (MR2206955) [10.1007/0-387-30528-9_8]
-\newblock L. Liberti,
-\newblock \emph{\emph{Writing global optimization software}},
-\newblock in \emph{Nonconvex Optimization and Its Applications}, Springer, \textbf{84} (2006), 211--262.
-
-\bibitem{linSchrage2009}% (MR2554904) [10.1080/10556780902753221]
-\newblock Y. Lin and L. Schrage,
-\newblock \emph{\emph{The global solver in the LINDO API}},
-\newblock \emph{Optim. Methods Softw.}, \textbf{24} (2009), 657--668.
-
 \bibitem{locatelliSchoenGO}% (MR3136805) [10.1137/1.9781611972672]
@@
 \newblock Springer-Verlag, Berlin, 2011.
-
-\bibitem{mongeauKarsentyRouze2000} %(MR1785196) [10.1080/10556780008805783]
-\newblock M. Mongeau, H. Karsenty, V. Rouz$\acute{e}$ and J. B. Hiriart-Urruty,
-\newblock \emph{\emph{Comparison of public-domain software for black box global optimization}},
-\newblock \emph{Optim. Methods Softw.}, \textbf{13} (2000), 203--226.
-
-\bibitem{mullen2014} %[10.18637/jss.v060.i06]
-\newblock K. M. Mullen,
-\newblock \emph{\emph{Continuous global optimization in R}},
-\newblock \emph{J. Stat. Softw.}, \textbf{60} (2014).
 
@@
 
-\bibitem{pinter2009}% (MR2528693)
-\newblock J. D. Pint\'er,
-\newblock \emph{\emph{Software development for global optimization}},
-\newblock \emph{Lectures on Global Optimization. Fields Institute Communications}, \textbf{55} (2009), 183--204.
-
-\bibitem{riosSahinidis2013} %(MR3070154) [10.1007/s10898-012-9951-y]
-\newblock L. M. Rios and N. V. Sahinidis,
-\newblock \emph{\emph{Derivative-free optimization: a review of algorithms and comparison of software implementations}},
-\newblock \emph{J. Glob. Optim.}, \textbf{56} (2013), 1247--1293.
-
-\bibitem{sahinidis1996} %(MR1376505) [10.1007/BF00138693]
-\newblock N. V. Sahinidis,
-\newblock \emph{\emph{BARON: A general purpose global optimization software package}},
-\newblock \emph{J. Glob. Optim.}, \textbf{8} (1996), 201--205.
-
-\bibitem{sergeyev1995} %(MR1358808) [10.1137/0805041]
-\newblock Y. D. Sergeyev,
-\newblock \emph{\emph{An information global optimization algorithm with local tuning}},
-\newblock \emph{SIAM J. Optim.}, \textbf{5} (1995), 858--870.
-
-\bibitem{sergeyev1999}% (MR1699326)
-\newblock Y. D. Sergeyev,
-\newblock \emph{\emph{Multidimensional global optimization using the first derivatives}},
-\newblock \emph{Comput. Math. Math. Phys.}, \textbf{39} (1999), 743--752.
-
-\bibitem{sergeyevKvasov2006}% (MR2197562) [10.1137/040621132]
-\newblock Y. D. Sergeyev and D. E. Kvasov,
-\newblock \emph{\emph{Global search based on efficient diagonal partitions and a set of Lipschitz constants}},
-\newblock \emph{SIAM Journal on Optimization}, \textbf{16} (2006), 910--937.
-
-\bibitem{sergeyevGrishagin2001}% (MR1825071) [10.1023/A:1010185125012]
-\newblock Y. D. Sergeyev and V. A. Grishagin,
-\newblock \emph{\emph{Parallel asynchronous global search and the nested optimization scheme}},
-\newblock \emph{J. Comput. Anal. Appl.}, \textbf{3} (2001), 123--145.
-
 \bibitem{sergeyevStronginLera2013}% (MR3113120) [10.1007/978-1-4614-8042-6]
@@
 \newblock Springer, 2013.
-
-\bibitem{Famularo2001}% (MR1866703) [10.1023/A:1012391611462]
-\newblock Y. D. Sergeyev, D. Famularo and P. Pugliese,
-\newblock \emph{\emph{Index branch-and-bound algorithm for Lipschitz univariate global optimization with multiextremal constraints}},
-\newblock \emph{J. Glob. Optim.}, \textbf{21} (2001), 317--341.
 
@@
 
-\bibitem{strongin1992}% (MR1263606) [10.1007/BF00122428]
+\bibitem{Strongin1992}% (MR1263606) [10.1007/BF00122428]
 \newblock R. G. Strongin,
@@
 
-\bibitem{venkataraman2009}
-\newblock P. Venkataraman,
-\newblock \emph{Applied Optimization with MATLAB Programming},
-\newblock John Wiley \& Sons, 2009.
-
 \bibitem{zhigljavskyRandGO}% (MR1187048) [10.1007/978-94-011-3436-1]
@@
 
-\bibitem{Two_Level_Parallel}
-Gergel, V., Sidorov, S.: A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems.
-In: Malyshkin, V.
-(Ed.) PaCT 2015, LNCS, vol.
-9251, pp.
-505-515.
-Springer, Heidelberg (2015)
-
 \bibitem{Strongin1991}
 Strongin, R.G.: Parallel multi-extremal optimization using a set of evolvents. Comp. Math. Math. Phys. \textbf{31(8)}, 37--46 (1991)
-
-\bibitem{Strongin1992}
-Strongin, R.G.: Algorithms for multi-extremal mathematical programming problems employing the set of joint space-filling curves. J. Global Optim. \textbf{2(4)}, 357--378 (1992)
 

### 2018-04-08 2018/RSD/MultiEvolvents/RSD.tex
RSD: drop unused references

@@
 \setlength{\floatsep}{4pt plus 1.0pt minus 1.5pt}
-%\setlength{\abovedisplayskip}{3pt}
-%\setlength{\belowdisplayskip}{3pt}
 
 \begin{document}
+\setlength{\abovedisplayskip}{3pt}
+\setlength{\belowdisplayskip}{3pt}
 %
@@
 \end{equation}
-where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbf{R}^N\) space.
+where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbb{R}^N\) space.
 \par
@@
 
-\bibitem{censorZeniosParGO}% (MR1486040)
-\newblock Y. Censor and S. A. Zenios,
-\newblock \emph{Parallel Optimization: Theory, Algorithms, and Applications},
-\newblock Oxford University Press, 1998.
-
-\bibitem{ciegisHentyParGO} (MR2499546)% [10.1007/978-0-387-09707-7]
-\newblock R. \v Ciegis, D. Henty, B. K\aa gstr\"om and J. \v Zilinskas,
-\newblock \emph{Parallel Scientific Computing and Optimization: Advances and Applications},
-\newblock Springer, 2009.
-
-\bibitem{Famularo1999} %(MR1827451) [10.1016/S0005-1098(99)00058-8]
-\newblock D. Famularo, P. Pugliese and Y. D. Sergeyev,
-\newblock \emph{\emph{A global optimization technique for checking parametric robustness}},
-\newblock \emph{Automatica}, \textbf{35} (1999), 1605--1611.
-
-\bibitem{fasanoPinter2013}% (MR3013800) [10.1007/978-1-4614-4469-5]
-\newblock G. Fasano and J. D. Pint\'er,
-\newblock \emph{Modeling and Optimization in Space Engineering},
-\newblock Springer, 2013.
-
-\bibitem{floudasPardalosGOState} %(MR1390521) [10.1007/978-1-4613-3437-8]
-\newblock C. A. Floudas and M. P. Pardalos,
-\newblock \emph{State of the Art in Global Optimization: Computational Methods and Applications},
-\newblock Kluwer Academic Publishers, Dordrecht, 1996.
-
-\bibitem{floudasPardGO}% (MR1147432) [10.1007/s10898-008-9332-8]
-\newblock C. A. Floudas and M. P. Pardalos,
-\newblock \emph{Recent Advances in Global Optimization},
-\newblock Princeton University Press, 2016.
-
-\bibitem{Gablonsky} %(MR1856800) [10.1023/A:1017930332101]
-\newblock J. M. Gablonsky and C. T. Kelley,
-\newblock \emph{\emph{A locally-biased form of the DIRECT algorithm}},
-\newblock \emph{J. Glob. Optim.}, \textbf{21} (2001), 27--37.
-
-\bibitem{horstTuyGO}% (MR1102239) [10.1007/978-3-662-03199-5]
-\newblock R. Horst and H. Tuy,
-\newblock \emph{Global Optimization: Deterministic Approaches},
-\newblock Springer-Verlag, Berlin, 1990.
-
-\bibitem{Kvasov2015}% (MR3585540) [10.1016/j.advengsoft.2014.09.014]
-\newblock D. E. Kvasov and Y. D. Sergeyev,
-\newblock \emph{\emph{Deterministic approaches for solving practical black-box global optimization problems}},
-\newblock \emph{Adv. Eng. Softw.}, \textbf{80} (2015), 58--66.
-
-\bibitem{Menniti}% [10.1016/j.epsr.2007.10.009]
-\newblock D. E. Kvasov, D. Menniti, A. Pinnarelli, Y. D. Sergeyev and N. Sorrentino,
-\newblock \emph{\emph{Tuning fuzzy power-system stabilizers in multi-machine systems by global optimization algorithms based on
- efficient domain partitions}},
-\newblock \emph{Electric Power Systems Research}, \textbf{78} (2008), 1217--1229.
-
-\bibitem{locatelliSchoenGO}% (MR3136805) [10.1137/1.9781611972672]
-\newblock M. Locatelli and F. Schoen,
-\newblock \emph{Global Optimization: Theory, Algorithms and Applications},
-\newblock SIAM, 2013.
-
-\bibitem{luqueAlbaGA}% (MR2808873) [10.1007/978-3-642-22084-5]
-\newblock G. Luque and E. Alba,
-\newblock \emph{Parallel Genetic Algorithms. Theory and Real World Applications},
-\newblock Springer-Verlag, Berlin, 2011.
-
-\bibitem{pardalosZhigljavskyZilinskas2016}% (MR2361744) [10.1007/978-3-319-29975-4]
-\newblock M. P. Pardalos, A. A. Zhigljavsky and J. \v Zilinskas,
-\newblock \emph{Advances in Stochastic and Deterministic Global Optimization},
-\newblock Springer, 2016.
-
-\bibitem{pinterGO}% (MR1374104) [10.1007/978-1-4757-2502-5]
-\newblock J. D. Pint\'er,
-\newblock \emph{Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications)},
-\newblock Kluwer Academic Publishers, Dordrecht, 1996.
-
 \bibitem{sergeyevStronginLera2013}% (MR3113120) [10.1007/978-1-4614-8042-6]

### 2018-04-09 2018/RSD/MultiEvolvents/RSD.tex
RSD: use square plots for evolvents
Добавлено введение и весьма краткое описание по гладким разверткам

@@
 \section{Introduction}\label{sec:intro}
+
+В данной статье рассматриваются параллельные алгоритмы решения задач многоэкстремальной оптимизации.
+В многоэкстремальных задачах возможность достоверной оценки глобального оптимума принципиально основана на наличии априорной информации о функции, позволяющей связать возможные значения оптимизируемой функции с известными значениями в точках осуществленных поисковых испытаний.
+Весьма часто такая информация о решаемой задаче представляется в виде предположения, что целевая функция $\varphi(y)$ удовлетворяет Lipschitz condition с неизвестной априори константой $L$ (см., например, \cite{Jones,Gablonsky,Evtushenko}).
+При этом целевая функция может быть задана программно, т.е.
+представлять собой “black-box”-функцию.
+Подобные задачи широко распространены в приложениях (problems of optimal design of objects and technological processes in various fields of technology, problems of model fitting according to observed data in scientific research, and so on).
+
+Многие методы, предназначенные для решения задач указанного класса, явно или неявно сводят решение многомерной задачи к решению одномерных подзадач (см., например, методы diagonal partitions \cite{Sergeyev2006,SergeyevKvasov2015} или simplicial partitions \cite{Zilinskas2008,Zilinskas2014}).
+В данной работе мы будем использовать подход, разработанный в ННГУ им.
+Н.И.
+Лобачевского и основанный на идее редукции размерности с помощью кривой Пеано $y(x)$, непрерывно и однозначно отображающей отрезок вещественной оси $[0,1]$ на $n$-мерный куб.
+
+В работах [ссылки] было предложено несколько способов построения разверток, аппроксимирующих теоретическую кривую Пеано.
+Указанные способы были реализованы в программной системе Globalizer.
+Целью настоящего исследования ставилось сравнения свойств разверток и выбор из них наиболее подходящей для использования в параллельных алгоритмах глобальной оптимизации.
 
@@
   \centering
-  \includegraphics[width=0.6\linewidth]{pictures/rotated_1.pdf}
+  \includegraphics[width=0.6\linewidth]{pictures/rotated.pdf}
   \caption{Two rotated evolvents on the same plane}
@@
 \subsection{Smooth evolvent}
+
+Рассмотренные в предыдущих пунктах способы построения развертки строят кривую $y(x)$, которая не является гладкой (см.
+рис.
+NN).
+Отсутствие гладкости может негативно сказаться на свойствах редуцированной одномерной функции $\varphi(y(x))$, т.к.
+гладкая кривая более качественно передает информации о возрастании/убывании исходной функции.
+На основе исходного алгоритма построения негладкой развертки было предложен обобщенный алгоритм, позволяющий строить гладкую space-filling.
+Для иллюстрации на рис.
+изображена гладкая развертка в двумерном случае.
 
@@
 
+\bibitem{Sergeyev2006}
+Sergeyev, Ya.D., Kvasov, D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim 16(3), 910–937 (2006)
+
+\bibitem{SergeyevKvasov2015} 
+Sergeyev, Y.D., Kvasov, D.E.: A deterministic global optimization using smooth diagonal auxiliary functions. Communications in Nonlinear Science and Numerical Simulation. 21(1-3), 99--111 (2015)
+
+\bibitem{Zilinskas2008}
+\v Zilinskas, J.: Branch and bound with simplicial partitions for global optimization. Math. Model. Anal.
+13(1), 145–159 (2008)
+
+\bibitem{Zilinskas2014}
+Paulavi\v cius, R., \v Zilinskas, J.: Simplicial Lipschitz optimization without the Lipschitz constant. J. Glob. Optim. 59(1), 23-40 (2014)
+
+\bibitem{Jones}
+Jones, D.R.: The direct global optimization algorithm. In: Floudas, C.A., Pardalos, P.M.
+(eds.) The Encyclopedia of Optimization, 2nd edn., pp. 725--735. Springer, Heidelberg
+(2009).
+
+\bibitem{Gablonsky}
+Gablonsky, J.M., Kelley, C.T.: A locally-biased form of the DIRECT algorithm. J. Glob. Optim., 21(1), 27--37 (2001)
+
+\bibitem{Evtushenko}
+Evtushenko, Y., Posypkin, M.: A deterministic approach to global box-constrained
+optimization. Optim. Lett. 7(4), 819--829 (2013)
+
 \end{thebibliography}

### 2018-04-10 2018/RSD/MultiEvolvents/RSD.tex
RSD: update plots and references

@@
 \section{Introduction}\label{sec:intro}
+\begin{Russian}
 
@@
 
-В работах [ссылки] было предложено несколько способов построения разверток, аппроксимирующих теоретическую кривую Пеано.
-Указанные способы были реализованы в программной системе Globalizer.
+В работах \cite{strongin1978,Strongin1992,Gergel2009,Goryachih2017} было предложено несколько способов построения разверток, аппроксимирующих теоретическую кривую Пеано.
+Указанные способы были реализованы в программной системе Globalizer \cite{globalizerSystem}.
 Целью настоящего исследования ставилось сравнения свойств разверток и выбор из них наиболее подходящей для использования в параллельных алгоритмах глобальной оптимизации.
 
+\end{Russian}
 %------------------------------------------------------------------------------
@@
 In this paper, the core class of optimization problems, which can be solved using
-Globalizer\cite{globalizerSystem}, is formulated. This class involves the multidimensional global
+Globalizer, is formulated. This class involves the multidimensional global
 optimization problems without constraints, which can be defined in the following way:
@@
 \subsection{Non injective evolvent}
+\begin{Russian}
 
@@
 На рис.
-\ref{fig:noninjective} изображена кривая с самопересечениями, полученная при соединении узлов грубой многомерной сетки в порядке следования их прообразов из отрезка $[0,1]$, а также отмечана точка, имеющая 3 прообраза.
+\ref{fig:noninjective} крестиками обозначена сетка в пространесве $\mathbb{R}^2$, для двух узлов которой указаны соответствующие им одномерное прообразы из $[0,1]$ (отмечены квадратами и кругами).
+Каждый указанный узел имеет по 3 прообраза.
 
@@
 рис.
-NN).
+\ref{fig:shifted_ev}).
 Отсутствие гладкости может негативно сказаться на свойствах редуцированной одномерной функции $\varphi(y(x))$, т.к.
 гладкая кривая более качественно передает информации о возрастании/убывании исходной функции.
-На основе исходного алгоритма построения негладкой развертки было предложен обобщенный алгоритм, позволяющий строить гладкую space-filling.
+На основе исходного алгоритма построения негладкой развертки было предложен обобщенный алгоритм \cite{Goryachih2017}, позволяющий строить гладкую space-filling curve.
 Для иллюстрации на рис.
-изображена гладкая развертка в двумерном случае.
+\ref{fig:smooth} изображена гладкая развертка в двумерном случае.
+Недостатком гладкой развёртки является в несколько раз большая сложность вычисления по сравнению с кусочно-линейными кривыми (требуется вычислять нелинейные гладкие функции).
+Причём с ростом точности аппроксимации и размерности количество интервалов гладкости увеличивается и сложность вычисления кривой нарастает.
+
+\end{Russian}
 
@@
 \section{Results of Numerical Experiments}
+\begin{Russian}
+
 The computational experiments have been carried out on the Lobachevsky supercomputer at
@@
 Параметры генератора для рассматриваемых классов приведены в \cite{Gaviano2003}.
+
+\end{Russian}
 
@@
 \label{sec:seq_comp}
+\begin{Russian}
 С целью понять, обладает ли какой-либо из перечисленных ранее типов развёрток существенным преимуществом над другими, были построены операционные характеристики индексного метода с различными типами развёрток на классах GKLS 2d Simple и GKLS 3d Simple.
@@
 Стоит заметить, что в расматриваемых задачах целевые функции не являются трудоёмкими (среднее время вычисления $10^{-3}$с) и в случае более сложных задач показатели ускорения по времени будут приближаться к ускорению по итерациям.
+\end{Russian}
 
@@
 \newblock Springer, 1989.
+
+\bibitem{Goryachih2017}
+\newblock Goryachih, A.
+\newblock A class of smooth modification of space-filling curves for global optimization problems
+\newblock Springer Proceedings in Mathematics and Statistics, 197, pp. 57-65, 2017
 

### 2018-04-16 2018/LaTeX/LeGo.tex

--- 2017-08-22 LION_Gergel/LION2017_gergel.tex
@@
-\documentclass{llncs}
+\documentclass{aip-cp}
 
-\usepackage{hyperref}
+\usepackage[numbers]{natbib}
+\usepackage{rotating}
 \usepackage{graphicx}
-\usepackage{multirow}
-\usepackage[misc,geometry]{ifsym}
+\pdfmapfile{+txfonts.map}
 
-\usepackage{amsmath}
-%\usepackage{amsfonts}
-%\usepackage{amssymb}
-%\usepackage{epstopdf}
-%\usepackage{epsfig}
-
-%\usepackage{url}
-%\urldef{\mailsa}\path|i.surmin@gmail.com, avbk@mail.ru, bastrakov@vmk.unn.ru,  |
-%\urldef{\mailsb}\path|evgeny.efimenko@gmail.com, arkady.gonoskov@gmail.com, meerov@vmk.unn.ru|
-
+% Document starts
 \begin{document}
 
-\mainmatter
+% Title portion
+\title{The Title Goes Here with Each Initial Letter Capitalized}
 
-\title{An Approach for Generating Test Problems of Constrained Global Optimization}
-\author{Victor Gergel \\
-\email{gergel@unn.ru}}
+\author{Victor Gergel}
+\author{Konstantin Barkalov\corref{cor1}}
+%\eaddress[url]{http://www.aip.org}
+\author{Ilya Lebedev}
+%\eaddress{anotherauthor@thisaddress.yyy}
 
-\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia}
+\affil{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia.}
+\corresp[cor1]{Corresponding author: konstantin.barkalov@itmm.unn.ru}
 
@@
 \begin{abstract}
-In the present paper, a novel approach to the constructing of the test global optimization problems with non-convex constraints is considered.
-The proposed approach is featured by a capability to construct the sets of such problems for carrying out multiple computational experiments in order to obtain a reliable evaluation of the efficiency of the optimization algorithms.
-When generating the test problems, the necessary number of constraints and desired fraction of the feasible domain relative to the whole global search domain can be specified.
-The locations of the global minimizers in the generated problems are known a priori that simplifies the evaluation of the results of the computational experiments essentially.
-A demonstration of the developed approach in the application to well-known index method for solving complex multiextremal optimization problems with non-convex constraints is presented.
-
-\keywords global optimization, multiextremal functions, non-convex constraints, test optimization problems, numerical experiments
-
+Abstract.
 \end{abstract}
 
-\section{Introduction}
+% Head 1
+\section{INTRODUCTION}
 
@@
 these ones may be defined by an algorithm for computing the function values in the search domain (so called ''black-box''-functions).
-It is supposed that even a single computing of a problem function value may be a time-consuming operation since it is related to the necessity of numerical modeling in the applied problems (see, for example, \cite{Famularo}--\cite{Modorskii}).
+It is supposed that even a single computing of a problem function value may be a time-consuming operation since it is related to the necessity of numerical modeling in the applied problems (see, for example, \cite{Famularo1999,Menniti2008}).
 
@@
 In order to obtain a reliable evaluation of the efficiency of the methods, the sets of test problems should be diverse and representative enough.
-The problem of choice of the test problems has been considered in a lot of works (see, for example, \cite{Floudas}--\cite{Addis}).
+The problem of choice of the test problems has been considered in a lot of works (see, for example, \cite{Floudas1999}).
 Unfortunately, in many cases, the proposed sets contain a small number of test problems, and it is difficult to obtain the problems with desired properties.
-And, the most important, the constraints are absent in the proposed test problems as a rule (or the constraints are relatively simple: linear, convex, etc.).
+The most important drawback consists of the fact that the constraints are absent in the proposed test problems as a rule (or the constraints are relatively simple: linear, convex, etc.).
 
-A novel approach to the generation of any amount of the global optimization problems with non-convex constraints for performing multiple computational experiments in order to obtain a reliable evaluation of the efficiency of the developed optimization algorithms has been proposed.
-When generating the test problems, the necessary number of constraints and desired fraction of the feasible domain relative to the whole search domain can be specified.
-In addition, the locations of the global minimizers in the generated problems are known a priori that simplifies the evaluation of the results of the computational experiments essentially.
+Do not abbreviate Figure , Equation, etc.; display items are always singular, i.e., Figure 1 and 2.
+Equations are always singular, i.e., Equation 1 and 2, and should be inserted using the Equation Editor, not as graphics, in the main text.
+Display items and captions should be inserted after the reference section.
+Please do not use footnotes in the text, additional information can be added to the reference list.
 
-\section{Test problem classes}
-A well-known approach to investigating and comparing the multiextremal optimization algorithms is based on testing these methods by solving a set of problems, chosen randomly from some specially designed class.
+% Acknowledgement
+\section{ACKNOWLEDGMENTS}
+This research was supported by the Russian Science Foundation, project No 16-11-10150.
 
-One generator for random samples of two-dimensional test functions has been described in \cite{Grishagin1978,Grishagin2016}. This generator produces two-dimensional functions according to the formula
-\begin{eqnarray} \nonumber \label{VAG}
-\varphi(y)= -&\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2+\right. \\
-&\left.\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},\\ \nonumber
-\end{eqnarray}
-where $g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2)$, $h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2)$, $y=(y_1,y_2)\in R^2, \; 0 \leq y_1,y_2 \leq 1$, and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
+% References
 
-Let us consider a scheme for constructing the generator GCGen (Global Constrained optimization problem Generator) which allows to generate the test global optimization problems with $m$ constraints.
-Obviously, one can generate $m+1$ functions, the first $m$ of these ones can be considered as the constraints and the $(m+1)$-th function -- as the objective one.
-However, in this case, the conditional global minimizer of the objective function is unknown, and the preliminary estimate of this one (for example, by scanning over a uniform grid) will be time-consuming.
-At the same time, one could not control the size of the feasible domain.
-In particular, the constraints might be incompatible, and the feasible domain might be empty.
-
-Below, the rules, which allow formulating the constrained global optimization problems so that:
-\begin{itemize}
-	\item one could control the size of feasible domain with respect to the whole domain of the parameters' variation;
-	\item the global minimizer of the objective function would be known a priori taking into account the constraints;
-	\item the global minimizer of the objective function without accounting for the constraints would be out of the feasible domain (with the purpose of simulating the behavior of the constraints and the objective function in the applied constrained optimization problems)
-\end{itemize}
-are proposed.
-
-The rules defining the operation of the generator of the constrained global optimization with the properties listed above consist in the following.
-
-\begin{enumerate}
-	\item Let us generate $m+1$ functions $f_j(y), y \in D, 1 \leq j \leq m+1,$ by some generating scheme (for instance, by using the formula (\ref{VAG})).
-The constraints will be constructed on the base of the first $m$ functions, the $(m+1)$-th function will serve for the construction of the objective function.
-	\item In order to know the global minimizer in the constrained problem a priori, let us make it to be the same to the global minimizer in the unconstrained problem.
-To do so, let us perform a linear transformation of coordinates so that the global minimizers of the constraint functions $y_j^\ast, \; 1 \leq j \leq m,$ would transit into the minimizer of the objective function $y_{m+1}^\ast$.
-This way, the functions $\overline{f_j}(y), \; 1 \leq j \leq m,$ with the same point of extremum will be constructed.
-		\item In order to control the size of the feasible domain, let us construct an auxiliary function (a combined constraint)
-		\[
-		H(y)=\max_{1 \leq j \leq m}{\overline{f_j}(y)}
-		\]
-and compute its values in the nodes of a uniform grid in the domain $D$; the number of the grid nodes in the conducted experiments should be big enough (in our experiments it was $\min \left\{10^7, 10^{2N} \right\}$).
-Then, let us find the maximum and minimum values of the function $H(y)$ in the grid nodes, $H_{max}$ and $H_{min}$, respectively, and construct a characteristic $s(i)$ -- the number of points, in which the values of $H(y)$ fall into the range
-\[
-\left[H_{min},H_{min}+i\frac{H_{max}-H_{min}}{100}\right], \; 1 \leq i \leq 100 .
-\]
-Then, the functions 
-\[
-\overline{f_j}(y) \leq q = H_{min}+i\frac{H_{max}-H_{min}}{100}, \; 1 \leq j \leq m,
-\]
-where $i$ is selected to be the minimal one satisfying the inequality
-\[
-\Delta \leq \frac{s(i)}{s(100)}
-\]
-will construct a problem with the feasible domain occupying the fraction $\Delta, \; 0<\Delta<1,$ of the whole search domain.
-\item 
-The test problem of global constrained optimization can be stated as follows
-\[
-\min\left\{ \varphi(y):y \in D, \; g_j(y)= \overline{f_j}(y)-q \leq 0, \; 1 \leq j \leq m \right\}
-\]
-where 
-%$\varphi(y)=f_{m+1}(y)$.
-This problem will possess all declared properties (the fraction of the feasible domain will consist $\Delta$ of the whole search domain and the global minimizer of the constrained problem will be known).
-%\item 
-%In order to complicate the test constrained problem the form of the objective function can be expanded by the expression
-\[
-\varphi(y) = f_{m+1}(y)-\beta\sum_{j=1}^m\max\left\{0,\overline{f_j}(y) -q \right\}^\alpha,
-\]
-where $\alpha$ is a positive integer number and $\beta>0$ is selected in such a way as to provide the global minimum location of the function $\varphi(y)$ in the infeasible part of the search domain $D$.
-For instance, to guarantee this property the value of $\beta$ can be set as follows
-\[
-\beta>(h_{max}-h_{min})/(H_{max}-q)^\alpha,
-\]
-where $h_{max}$ and $h_{min}$ are the maximum and minimum values of the function $f_{m+1}(y)$ respectively.
-
-\end{enumerate}
-
-\section{Some numerical results}
-
-As an illustration, the level lines of the objective functions and the zero-level lines of the constraints for problems constructed on the base of functions (\ref{VAG}) with $\alpha = 3, \beta = 1$ and the volume fractions of the feasible domains $\Delta = 0.4, 0.6$ are shown in Fig.~\ref{fig:VAG}.
-The feasible domains are highlighted by green.
-The change of volume and, at the same time, the increase of complexity of the feasible domains are seen clearly.
-Fig.~\ref{fig:VAG} (a,b) also shows the points of 628 and 764 trials, correspondingly, performed by the \textit{index method} for solving constrained global optimization problems until the required accuracy $\epsilon=10^{-2}$ was achieved.
-The conditional optimizer is shown as a red point and the best estimation of the optimizer is shown as a blue point.
-
-The index method has been proposed and developed in \cite{Strongin2000,Barkalov2002}.
-The approach is based on a separate accounting for each constraint of the problem and is not related to the use of the penalty functions.
-According to the rules of index method, every iteration includes a sequential checking of fulfillment of the problem constraints at this point.
-The first occurrence of violation of any constraint terminates the trial and initiates the transition to the next iteration.
-This allows: (i) accounting for the information on each constraint separately and (ii) solving the problems, in which the function values may be undefined out of the feasible domain.
-It should be noted, that the index method can be efficiently parallelized for accelerators \cite{Barkalov2015,Barkalov2016}.
-
-\begin{figure}
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{vag_04.png} \\ (a)}
-\end{minipage}
-\hfill
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{vag_06.png} \\ (b)}
-\end{minipage}
-\caption{The problems based on functions (\ref{VAG}) }
-\label{fig:VAG}
-\end{figure}
-
-\section {Conclusion}
-
-This paper considers the method for generating global optimization test problems with non-convex constraints that allows:
-\begin{itemize}
-	\item to control the size of feasible domain with respect to the whole domain of the parameters' variation;
-	\item to known a priori the conditional global minimizer of the objective function;
-	\item to generate the unconditional global minimizer of the objective function out of the feasible domain (to simulate the constraints and objective function in the applied optimization problems).
-\end{itemize}
-
-The demonstration of the developed approach in application to well-known index method for solving complex multiextremal optimization problems with non-convex constraints is considered.
-
-The developed approach allows generating any number of test global optimization problems with non-convex constraints for performing multiple computational experiments in order to obtain a reliable evaluation of the efficiency of the developed optimization algorithms.
-To develop the proposed approach further, the development of new test classes for the optimization problems of various dimensionalities is planned.
-
-\bigskip
-
-\textbf{Acknowledgments.} This study was supported by the Russian Science Foundation, project No 16-11-10150.
-
-\begin{thebibliography}{99}
-
-\bibitem{Famularo}
-Famularo, D., Pugliese, P., Sergeyev, Ya.D.:A global optimization technique for checking parametric robustness. Automatica. 35, 1605--1611 (1999)
-
-\bibitem{Menniti}
-Kvasov, D.E., Menniti, D., Pinnarelli, A., Sergeyev, Ya.D., Sorrentino, N.: Tuning fuzzy power-system stabilizers in multi-machine systems by global optimization algorithms based on efficient domain partitions.
-Electric Power Systems Research.
-78(7), 1217--1229 (2008)
-
-\bibitem{Kvasov2015}
-Kvasov, D.E., Sergeyev, Y.D.: Deterministic approaches for solving practical black-box global optimization problems. Advances in Engineering Software. 80, 58--66 (2015)
-
-\bibitem{Modorskii}
-Modorskii, V.Y., Gaynutdinova, D.F., Gergel, V.P., Barkalov, K.A.: Optimization in design of scientific products for purposes of cavitation problems.Solving Global Optimization Problems on GPU Cluster.
-In: Simos T.E.
-(Ed.) ICNAAM 2015, AIP Conference Proceedings.
-1738, art.
-no.
-400013 (2016)
-
-\bibitem{Floudas} 
-Floudas, C.A., et al.: Handbook of test problems in local and global optimization. Kluwer Academic Publishers, Dordrecht (1999)
-
-\bibitem{Gaviano}
-Gaviano, M., Kvasov, D.E., Lera, D., Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM TOMS 29(4), 469--480 (2003)
-
-\bibitem{Ali} Ali, M.M., Khompatraporn, C., Zabinsky Z.B.: A numerical evaluation of several stochastic algorithms on selected continuous global optimization test problems.
-J.
-Glob.
-Optim.
-31(4), 635--672 (2005)
-
-\bibitem{Addis} Addis, B., Locatelli, M.: A new class of test functions for global optimization. J. Glob. Optim. 38(3), 479--501 (2007)
-
-\bibitem{Grishagin1978}
-Grishagin, V.A.: Operating Characteristics of Some Global Search Algorithms. Problems of Statistical Optimization. 7, 198--206 (1978) [in Russian]
-
-\bibitem{Grishagin2016} 
-Gergel, V., Grishagin, V., Gergel, A.: Adaptive nested optimization scheme for multidimensional global search. J. Glob. Optim. 66 (1), 35--51 (2016)
-
-\bibitem{Strongin2000}
-Strongin, R.G., Sergeyev, Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000)
-
-\bibitem{Pugliese}
-Sergeyev, Ya.D., Famularo, D., Pugliese, P.: Index Branch-and-Bound Algorithm for Lipschitz univariate global optimization with multiextremal constraints. J. Glob. Optim. 21(3), 317--341 (2001)
-
-\bibitem{Barkalov2002}
-Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)
-
-%\bibitem{Strongin2003} Strongin, R.G., Sergeyev, Ya.D.: Global optimization: fractal approach and non-redundant parallelism. J. Glob. Optim. 27(1), 25--50 (2003)
-
-%\bibitem{Barkalov2010} Barkalov, K., Ryabov, V., Sidorov, S.: Parallel Scalable Algorithms with Mixed Local-Global Strategy for Global Optimization Problems.
-In: Malyshkin, V.
-(Ed.) MTPP 2010, LNCS.
-6083, 232--240 (2010)
-
-%\bibitem{Barkalov2014} Barkalov, K.A., Gergel, V.P.: Multilevel scheme of dimensionality reduction for parallel global search algorithms.
-In: Proceedings of the 1st International Conference on Engineering and Applied Sciences Optimization -- OPT-i 2014.
-pp.
-2111--2124 (2014)
-
-\bibitem{Barkalov2015}
-Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi Coprocessor for Solving Global Optimization Problems. In: Malyshkin, V. (Ed.) PaCT 2015, LNCS. 9251, 307--318 (2015)
-
-\bibitem{Barkalov2016}
-Barkalov, K., Gergel, V.: Parallel Global Optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016)
-
-%\bibitem{Lebedev2016}Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster.
-In: Simos T.E.
-(Ed.) ICNAAM 2015, AIP Conference Proceedings.
-1738, art.
-no.
-400006 (2016)
-
-\end{thebibliography}
+\nocite{*}
+\bibliographystyle{aipnum-cp}%
+\bibliography{LeGo}%
 

### 2018-04-16 2018/LeGo/LeGo.tex
Paper from LION

--- 2017-08-22 LION_Gergel/LION2017_gergel.tex
@@
 The analytical formulae of the problem functions may be unknown, i.e.
-these ones may be defined by an algorithm for computing the function values in the search domain (so called ''black-box''-functions).
-It is supposed that even a single computing of a problem function value may be a time-consuming operation since it is related to the necessity of numerical modeling in the applied problems (see, for example, \cite{Famularo}--\cite{Modorskii}).
+these ones may be defined by an algorithm for computing the function values in the search domain (so called ``black-box''-functions).
+It is supposed that even a single computing of a problem function value may be a time-consuming operation since it is related to the necessity of numerical modeling in the applied problems (see, for example, \cite{Menniti2008,Kvasov2015,Modorskii2017}).
 
@@
 In order to obtain a reliable evaluation of the efficiency of the methods, the sets of test problems should be diverse and representative enough.
-The problem of choice of the test problems has been considered in a lot of works (see, for example, \cite{Floudas}--\cite{Addis}).
+The problem of choice of the test problems has been considered in a lot of works (see, for example, \cite{Floudas1999,Gaviano2003,Ali2005,Addis2007}).
 Unfortunately, in many cases, the proposed sets contain a small number of test problems, and it is difficult to obtain the problems with desired properties.
-And, the most important, the constraints are absent in the proposed test problems as a rule (or the constraints are relatively simple: linear, convex, etc.).
-
-A novel approach to the generation of any amount of the global optimization problems with non-convex constraints for performing multiple computational experiments in order to obtain a reliable evaluation of the efficiency of the developed optimization algorithms has been proposed.
-When generating the test problems, the necessary number of constraints and desired fraction of the feasible domain relative to the whole search domain can be specified.
-In addition, the locations of the global minimizers in the generated problems are known a priori that simplifies the evaluation of the results of the computational experiments essentially.
+The most important drawback consists of the fact that the constraints are absent in the proposed test problems as a rule (or the constraints are relatively simple: linear, convex, etc.).
 
@@
 
-One generator for random samples of two-dimensional test functions has been described in \cite{Grishagin1978,Grishagin2016}. This generator produces two-dimensional functions according to the formula
-\begin{eqnarray} \nonumber \label{VAG}
-\varphi(y)= -&\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2+\right. \\
-&\left.\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},\\ \nonumber
-\end{eqnarray}
-where $g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2)$, $h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2)$, $y=(y_1,y_2)\in R^2, \; 0 \leq y_1,y_2 \leq 1$, and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
+%One generator for random samples of two-dimensional test functions has been described in \cite{Grishagin1978,Grishagin2016}. This generator produces two-dimensional functions according to the formula
+%\begin{eqnarray} \nonumber \label{VAG}
+%\varphi(y)= -&\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2+\right. \\
+%&\left.\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},\\ \nonumber
+%\end{eqnarray}
+%where $g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2)$, $h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2)$, $y=(y_1,y_2)\in R^2, \; 0 \leq y_1,y_2 \leq 1$, and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
 
@@
 
-\section{Some numerical results}
+\section{Numerical results}
 
@@
 
-The index method has been proposed and developed in \cite{Strongin2000,Barkalov2002}.
+The index method has been proposed and developed in \cite{Strongin2000,Sergeyev2001,Barkalov2002}.
 The approach is based on a separate accounting for each constraint of the problem and is not related to the use of the penalty functions.
@@
 The developed approach allows generating any number of test global optimization problems with non-convex constraints for performing multiple computational experiments in order to obtain a reliable evaluation of the efficiency of the developed optimization algorithms.
-To develop the proposed approach further, the development of new test classes for the optimization problems of various dimensionalities is planned.
 
-\bigskip
+% Acknowledgement
+\section{ACKNOWLEDGMENTS}
+This research was supported by the Russian Science Foundation, project No 16-11-10150.
 
-\textbf{Acknowledgments.} This study was supported by the Russian Science Foundation, project No 16-11-10150.
+% References
 
-\begin{thebibliography}{99}
-
-\bibitem{Famularo}
-Famularo, D., Pugliese, P., Sergeyev, Ya.D.:A global optimization technique for checking parametric robustness. Automatica. 35, 1605--1611 (1999)
-
-\bibitem{Menniti}
-Kvasov, D.E., Menniti, D., Pinnarelli, A., Sergeyev, Ya.D., Sorrentino, N.: Tuning fuzzy power-system stabilizers in multi-machine systems by global optimization algorithms based on efficient domain partitions.
-Electric Power Systems Research.
-78(7), 1217--1229 (2008)
-
-\bibitem{Kvasov2015}
-Kvasov, D.E., Sergeyev, Y.D.: Deterministic approaches for solving practical black-box global optimization problems. Advances in Engineering Software. 80, 58--66 (2015)
-
-\bibitem{Modorskii}
-Modorskii, V.Y., Gaynutdinova, D.F., Gergel, V.P., Barkalov, K.A.: Optimization in design of scientific products for purposes of cavitation problems.Solving Global Optimization Problems on GPU Cluster.
-In: Simos T.E.
-(Ed.) ICNAAM 2015, AIP Conference Proceedings.
-1738, art.
-no.
-400013 (2016)
-
-\bibitem{Floudas} 
-Floudas, C.A., et al.: Handbook of test problems in local and global optimization. Kluwer Academic Publishers, Dordrecht (1999)
-
-\bibitem{Gaviano}
-Gaviano, M., Kvasov, D.E., Lera, D., Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM TOMS 29(4), 469--480 (2003)
-
-\bibitem{Ali} Ali, M.M., Khompatraporn, C., Zabinsky Z.B.: A numerical evaluation of several stochastic algorithms on selected continuous global optimization test problems.
-J.
-Glob.
-Optim.
-31(4), 635--672 (2005)
-
-\bibitem{Addis} Addis, B., Locatelli, M.: A new class of test functions for global optimization. J. Glob. Optim. 38(3), 479--501 (2007)
-
-\bibitem{Grishagin1978}
-Grishagin, V.A.: Operating Characteristics of Some Global Search Algorithms. Problems of Statistical Optimization. 7, 198--206 (1978) [in Russian]
-
-\bibitem{Grishagin2016} 
-Gergel, V., Grishagin, V., Gergel, A.: Adaptive nested optimization scheme for multidimensional global search. J. Glob. Optim. 66 (1), 35--51 (2016)
-
-\bibitem{Strongin2000}
-Strongin, R.G., Sergeyev, Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000)
-
-\bibitem{Pugliese}
-Sergeyev, Ya.D., Famularo, D., Pugliese, P.: Index Branch-and-Bound Algorithm for Lipschitz univariate global optimization with multiextremal constraints. J. Glob. Optim. 21(3), 317--341 (2001)
-
-\bibitem{Barkalov2002}
-Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)
-
-%\bibitem{Strongin2003} Strongin, R.G., Sergeyev, Ya.D.: Global optimization: fractal approach and non-redundant parallelism. J. Glob. Optim. 27(1), 25--50 (2003)
-
-%\bibitem{Barkalov2010} Barkalov, K., Ryabov, V., Sidorov, S.: Parallel Scalable Algorithms with Mixed Local-Global Strategy for Global Optimization Problems.
-In: Malyshkin, V.
-(Ed.) MTPP 2010, LNCS.
-6083, 232--240 (2010)
-
-%\bibitem{Barkalov2014} Barkalov, K.A., Gergel, V.P.: Multilevel scheme of dimensionality reduction for parallel global search algorithms.
-In: Proceedings of the 1st International Conference on Engineering and Applied Sciences Optimization -- OPT-i 2014.
-pp.
-2111--2124 (2014)
-
-\bibitem{Barkalov2015}
-Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi Coprocessor for Solving Global Optimization Problems. In: Malyshkin, V. (Ed.) PaCT 2015, LNCS. 9251, 307--318 (2015)
-
-\bibitem{Barkalov2016}
-Barkalov, K., Gergel, V.: Parallel Global Optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016)
-
-%\bibitem{Lebedev2016}Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster.
-In: Simos T.E.
-(Ed.) ICNAAM 2015, AIP Conference Proceedings.
-1738, art.
-no.
-400006 (2016)
-
-\end{thebibliography}
+%\nocite{*}
+\bibliographystyle{aipnum-cp}%
+\bibliography{LeGo}%
 

### 2018-04-19 2018/LION_Mixed-Integer/LION2018.tex
Camera-ready version

@@
 
-As an illustration, in Fig. \ref{fig:1} presents the plots of the functions corresponding to a 
+As an illustration, Fig. \ref{fig:1} presents the plots of the functions corresponding to a 
 problem with one continuous parameter and one binary parameter.
@@
 the subproblem, the solving of which corresponds to the solving of the initial 
-problem (\ref{problem_i}). In the rest subproblems, a minor part of trials only will be performed 
+problem (\ref{problem_i}). In the rest subproblems, a only minor part of trials will be performed 
 since the solutions of these subproblems are the locally optimal ones.
@@
 Thus, we have constructed the \textit{Mixed-Integer Global Search Algorithm} (MIGSA) based on the  
-reduction of the MINLP to the NLP. 
+reduction of the mixed-integer non-convex optimization problem to the non-convex optimization problem. 
 %It is possible to prove that the convergence conditions for this algorithm will follow from the convergence conditions of its prototype (GSA).
@@
 
-Let us compare the proposed MIGSA method with a genetic algorithm for solving the mixed-integer optimization problems implemented in Matlab Global Optimization Toolbox. In Table 
+Let us compare the proposed MIGSA method with a genetic algorithm for solving the mixed-integer optimization problems implemented in Matlab Global Optimization Toolbox \cite{Matlab}. In Table 
 \ref{tab:1}, the numbers of trials required for solving the known test mixed-integer 
@@
 \bibitem{Strongin2000}
-Strongin, R.G., Sergeyev, Y.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000) %; DOI: 10.1007/978-1-4615-4677-1
+Strongin, R.G., Sergeyev, Y.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000) %; DOI: 10.1007/978-1-4615-4677-1
 
@@
 \bibitem{Floudas}
-Floudas, C.A., Pardalos, P.M.:  Handbook of Test Problems in Local and Global Optimization. Springer (1999)  %; DOI: 10.1007/978-1-4757-3040-1
+Floudas, C.A., Pardalos, P.M.:  Handbook of test problems in local and global optimization. Springer (1999)  %; DOI: 10.1007/978-1-4757-3040-1
+
+\bibitem{Matlab}
+https://www.mathworks.com/help/gads/mixed-integer-optimization.html
 
@@
 \bibitem{Lebedev2015}
-Lebedev, I., Gergel, V.: Heterogeneous Parallel Computations for Solving Global Optimization Problems. Procedia Computer Science \textbf{66}, 53--62 (2015)
+Lebedev, I., Gergel, V.: Heterogeneous parallel computations for solving global optimization problems. Procedia Computer Science \textbf{66}, 53--62 (2015)
 
 \bibitem{Gergel2015}
-Gergel, V., Sidorov, S.: A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems.
+Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of computationally intensive multiextremal optimization problems.
 Lecture Notes in Computer Science  \textbf{9251}, 505--515 (2015)

### 2018-04-20 2018/LeGo/LeGo.tex
Russian text

@@
 \pdfmapfile{+txfonts.map}
+
+%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
+\usepackage[utf8x]{inputenc}
+\usepackage[english,russian]{babel}
+\usepackage{cmap}
+%%%%%
 
@@
 where $h_{max}$ and $h_{min}$ are the maximum and minimum values of the function $f_{m+1}(y)$ respectively.
+\end{enumerate}
 
+\Russian
+Одним из важных свойств, характеризующих прикладные задачи условной оптимизации, является расположение условного минимума на границе feasible domain.
+При генерации задач по схеме, описанной в \cite{Gergel2017}, условный минимум будет, вообще говоря, находиться внутри feasible domain.
+Поэтому в новой версиии генератора задач была добавлена возможность как смещения условного минимума на границу, так и выбора числа активных ограничений в нем.
+
+Смещение условного минимума на границу feasible domain осуществляется следующим образом.
+\begin{enumerate}
+	\item Генерируется задача с произвольным расположением условного минимума $\overline{y}$;
+	\item Из точки $\overline{y}$ с заданным шагом $h$ осуществляется покоординатный поиск допустимой точки $y^*$, для которой хотя бы одно из ограничений является активным, т.е.
+существует такой номер $j, 1 \leq j\leq m$, что $\left|g_j(y^*)\right| \leq \delta$.
+	\item Полученное грубое значение $y^*$ уточняется локальным методом. 
+	\item Выполняется преобразование координат, переводящее точку $\overline{y}$ в точку $y^*$. Тем самым целевая функция достигает своего минимума на границе feasible domain. 
+\end{enumerate}
+
+Отметим, что так как поиск точки $y^*$ осуществляется по каждой координате в отдельности, то трудоемкость поиска будет возрастать линейно при увеличении размерности задачи.
+
+В случае, если нужно сгенерировать задачу с заданным числом $S$ активных ограничений в точке условного минимума, то тогда правило 2 меняется следующим образом 
+\begin{enumerate}
+	\item В области поиска строится uniform grid с числом точек $\min \left\{10^7, 10^{2N} \right\}$, в каждой из которых вычисляются значения ограничений. 
+	\item Определяется множество точек $\{y^t\}$, в которых активными являются $S$ ограничений; т.е.
+существуют такие номера $j_1, ...,j_S$, что $\left|g_{j_s}(y^t)\right|\leq \delta, \; 1\leq s \leq S$.
+В случае отсутствия такой точки, генерируется задача с произвольным числом активных ограничений и пользователь информируется об этом.
+	\item Выбирается точка $y^* = \min_{t} \{\left\|\overline{y} - y^t\right\|\}$. Полученное грубое значение $y^*$ уточняется локальным методом. 
 \end{enumerate}

### 2018-04-23 2018/LeGo/LeGo.tex

@@
 \usepackage{graphicx}
+\usepackage{enumerate}
 \pdfmapfile{+txfonts.map}
@@
 
-\author{Victor Gergel}
-\author{Konstantin Barkalov\corref{cor1}}
+\author{Victor Gergel\corref{cor1}}
+\author{Konstantin Barkalov}
 %\eaddress[url]{http://www.aip.org}
@@
 %\eaddress{anotherauthor@thisaddress.yyy}
+\author{Maria Rachinskaya}
 
 \affil{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia.}
-\corresp[cor1]{Corresponding author: konstantin.barkalov@itmm.unn.ru}
+\corresp[cor1]{Corresponding author: gergel@unn.ru}
 
@@
 \begin{abstract}
-Abstract.
+В работе рассматривается дальнейшее развитие подхода к построению тестовых задач глобальной оптимизации с нелинейными ограничениями.
+Для порождаемых задач расположение глобального минимума является известным.
+Отличительной чертой рассматриваемого генератора является возможность указания желаемого количества ограничений и доли допустимой области относительно все области глобального поиска.
+В обновленной версии генератора добавлена возможность формирования задач с условным минимумом, расположенным на границе допустимой области, и указания числа ограничений, активных в точке минимума.
+Демонстрация разработанного подхода приводится на примере известного индексного метода для решения сложных многоэкстремальных задач оптимизации с нелинейными ограничениями.
 \end{abstract}
@@
 
+В данной работе проводится дальнейшее развитие подхода для порождения задач глобальной оптимизации с нелинейными ограничениями, первоначально предложенного в \cite{Gergel2017}.
+При генерации тестовых задач может быть указано необходимое количество ограничений и желаемая доля допустимой области относительно все области глобального поиска.
+При этом для порождаемых задач расположение глобального минимума является известным, что существенно упрощает оценку результатов выполненных вычислительных экспериментов.
+В очередную версию генератора, в дополнение к указанным свойствам, были добавлены следующие возможности.
+\begin{itemize}
+	\item Возможность генерации задач с точкой условного минимума, расположенной на границе допустимой области.
+	\item Возможность указания числа ограничений, активных в точке оптимума.
+\end{itemize}
+Указанные возможности позволяют на более высоком уровне имитировать свойства прикладных задач условной глобальной оптимизации.
+
 \section{Test problem classes}
@@
 
-%One generator for random samples of two-dimensional test functions has been described in \cite{Grishagin1978,Grishagin2016}. This generator produces two-dimensional functions according to the formula
-%\begin{eqnarray} \nonumber \label{VAG}
-%\varphi(y)= -&\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2+\right. \\
-%&\left.\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},\\ \nonumber
-%\end{eqnarray}
-%where $g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2)$, $h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2)$, $y=(y_1,y_2)\in R^2, \; 0 \leq y_1,y_2 \leq 1$, and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
+For example, one generator for random samples of two-dimensional test functions has been described in \cite{Grishagin2015}. This generator produces two-dimensional functions according to the formula
+\begin{equation} \label{VAG}
+\varphi(y)= -\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2 + \left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},
+\end{equation}
+where $g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2)$, $h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2)$, $y=(y_1,y_2)\in R^2, \; 0 \leq y_1,y_2 \leq 1$, and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
 
-Let us consider a scheme for constructing the generator GCGen (Global Constrained optimization problem Generator) which allows to generate the test global optimization problems with $m$ constraints.
-Obviously, one can generate $m+1$ functions, the first $m$ of these ones can be considered as the constraints and the $(m+1)$-th function -- as the objective one.
-However, in this case, the conditional global minimizer of the objective function is unknown, and the preliminary estimate of this one (for example, by scanning over a uniform grid) will be time-consuming.
-At the same time, one could not control the size of the feasible domain.
-In particular, the constraints might be incompatible, and the feasible domain might be empty.
+Another generator (\textit{GKLS generator}) for the functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.) has been proposed in \cite{Gaviano2003}.
+Application of this generator for studying some unconstrained optimization algorithms has been described in \cite{Paulavicius2014,Sergeyev2015}.
 
-Below, the rules, which allow formulating the constrained global optimization problems so that:
+The generator GCGen (Global Constrained optimization problem Generator) which allows to generate the test global optimization problems with $m$ constraints на основе функций заданного класса, has been proposed in \cite{Gergel2017}.
+В данной статье были были предложены the rules, which allow formulating the constrained global optimization problems so that:
 \begin{itemize}
@@
 	\item the global minimizer of the objective function would be known a priori taking into account the constraints;
-	\item the global minimizer of the objective function without accounting for the constraints would be out of the feasible domain (with the purpose of simulating the behavior of the constraints and the objective function in the applied constrained optimization problems)
+	\item the global minimizer of the objective function without accounting for the constraints would be out of the feasible domain.
 \end{itemize}
-are proposed.
-
-The rules defining the operation of the generator of the constrained global optimization with the properties listed above consist in the following.
-
-\begin{enumerate}
-	\item Let us generate $m+1$ functions $f_j(y), y \in D, 1 \leq j \leq m+1,$ by some generating scheme (for instance, by using the formula (\ref{VAG})).
-The constraints will be constructed on the base of the first $m$ functions, the $(m+1)$-th function will serve for the construction of the objective function.
-	\item In order to know the global minimizer in the constrained problem a priori, let us make it to be the same to the global minimizer in the unconstrained problem.
-To do so, let us perform a linear transformation of coordinates so that the global minimizers of the constraint functions $y_j^\ast, \; 1 \leq j \leq m,$ would transit into the minimizer of the objective function $y_{m+1}^\ast$.
-This way, the functions $\overline{f_j}(y), \; 1 \leq j \leq m,$ with the same point of extremum will be constructed.
-		\item In order to control the size of the feasible domain, let us construct an auxiliary function (a combined constraint)
-		\[
-		H(y)=\max_{1 \leq j \leq m}{\overline{f_j}(y)}
-		\]
-and compute its values in the nodes of a uniform grid in the domain $D$; the number of the grid nodes in the conducted experiments should be big enough (in our experiments it was $\min \left\{10^7, 10^{2N} \right\}$).
-Then, let us find the maximum and minimum values of the function $H(y)$ in the grid nodes, $H_{max}$ and $H_{min}$, respectively, and construct a characteristic $s(i)$ -- the number of points, in which the values of $H(y)$ fall into the range
-\[
-\left[H_{min},H_{min}+i\frac{H_{max}-H_{min}}{100}\right], \; 1 \leq i \leq 100 .
-\]
-Then, the functions 
-\[
-\overline{f_j}(y) \leq q = H_{min}+i\frac{H_{max}-H_{min}}{100}, \; 1 \leq j \leq m,
-\]
-where $i$ is selected to be the minimal one satisfying the inequality
-\[
-\Delta \leq \frac{s(i)}{s(100)}
-\]
-will construct a problem with the feasible domain occupying the fraction $\Delta, \; 0<\Delta<1,$ of the whole search domain.
-\item 
-The test problem of global constrained optimization can be stated as follows
-\[
-\min\left\{ \varphi(y):y \in D, \; g_j(y)= \overline{f_j}(y)-q \leq 0, \; 1 \leq j \leq m \right\}
-\]
-where 
-%$\varphi(y)=f_{m+1}(y)$.
-This problem will possess all declared properties (the fraction of the feasible domain will consist $\Delta$ of the whole search domain and the global minimizer of the constrained problem will be known).
-%\item 
-%In order to complicate the test constrained problem the form of the objective function can be expanded by the expression
-\[
-\varphi(y) = f_{m+1}(y)-\beta\sum_{j=1}^m\max\left\{0,\overline{f_j}(y) -q \right\}^\alpha,
-\]
-where $\alpha$ is a positive integer number and $\beta>0$ is selected in such a way as to provide the global minimum location of the function $\varphi(y)$ in the infeasible part of the search domain $D$.
-For instance, to guarantee this property the value of $\beta$ can be set as follows
-\[
-\beta>(h_{max}-h_{min})/(H_{max}-q)^\alpha,
-\]
-where $h_{max}$ and $h_{min}$ are the maximum and minimum values of the function $f_{m+1}(y)$ respectively.
-\end{enumerate}
 
@@
 
-В случае, если нужно сгенерировать задачу с заданным числом $S$ активных ограничений в точке условного минимума, то тогда правило 2 меняется следующим образом 
+В случае, если нужно сгенерировать задачу с заданным числом $S$ активных ограничений в точке условного минимума, то тогда правила меняются следующим образом 
 \begin{enumerate}
-	\item В области поиска строится uniform grid с числом точек $\min \left\{10^7, 10^{2N} \right\}$, в каждой из которых вычисляются значения ограничений. 
-	\item Определяется множество точек $\{y^t\}$, в которых активными являются $S$ ограничений; т.е.
-существуют такие номера $j_1, ...,j_S$, что $\left|g_{j_s}(y^t)\right|\leq \delta, \; 1\leq s \leq S$.
-В случае отсутствия такой точки, генерируется задача с произвольным числом активных ограничений и пользователь информируется об этом.
-	\item Выбирается точка $y^* = \min_{t} \{\left\|\overline{y} - y^t\right\|\}$. Полученное грубое значение $y^*$ уточняется локальным методом. 
+	\item Генерируется задача с произвольным расположением условного минимума $\overline{y}$;
+	\item Из точки $\overline{y}$ с заданным шагом $h$ осуществляется покоординатный поиск допустимой точки $y^*$, для которой хотя бы одно из ограничений является активным, т.е.
+существует такой номер $j, 1 \leq j\leq m$, что $\left|g_j(y^*)\right| \leq \delta$.
+	\item Полученное грубое значение $y^*$ уточняется локальным методом. 
+	\item Определяется число $K$ ограничений, активных в точке $y^*$.
+	\item Если $K>S$, то из числа активных выбираются $K-S$ ограничений, и путем вычитания положительного параметра $q$ из правой части они переводятся в неактивные.
+	\item Если $K<S$, то из числа неактивных выбираются $S-K$ ограничений, и путем прибавления положительного параметра $q$ в правой части они переводятся в активные.
+	\item Выполняется преобразование координат, переводящее точку $\overline{y}$ в точку $y^*$.
+Тем самым целевая функция достигает своего минимума на границе feasible domain с заданным числом активных ограничений.
 \end{enumerate}
@@
 
-As an illustration, the level lines of the objective functions and the zero-level lines of the constraints for problems constructed on the base of functions (\ref{VAG}) with $\alpha = 3, \beta = 1$ and the volume fractions of the feasible domains $\Delta = 0.4, 0.6$ are shown in Fig.~\ref{fig:VAG}.
+As an illustration, the level lines of the objective functions and the zero-level lines of four constraints for the problem constructed on the base of functions (\ref{VAG}) with the volume fraction of the feasible domains $\Delta = 0.5$ are shown in Fig.~\ref{fig:VAG}.
 The feasible domains are highlighted by green.
-The change of volume and, at the same time, the increase of complexity of the feasible domains are seen clearly.
-Fig.~\ref{fig:VAG} (a,b) also shows the points of 628 and 764 trials, correspondingly, performed by the \textit{index method} for solving constrained global optimization problems until the required accuracy $\epsilon=10^{-2}$ was achieved.
+
+Fig.~\ref{fig:VAG} (a,b,c,d) соответствуют разным способам генерации тестовых задач.
+\begin{description}
+	\item[(a)] Задача сгенерирована с расположением constrained optimizer внутри feasible domain.
+	\item[(b)] Модификация задачи из п. (a): unconstrained optimizer находится вне feasible domain.
+	\item[(c)] Модификация задачи из п. (b): constrained optimizer находится на границе feasible domain, число активных ограничений не задано.
+	\item[(d)] Модификация задачи из п. (c): число активных ограничений задано равным двум.
+\end{description}
+
+Fig.~\ref{fig:VAG} (a,b,c,d) also shows the points of 628 and 764 trials, correspondingly, performed by the \textit{index method} for solving constrained global optimization problems until the required accuracy $\epsilon=10^{-2}$ was achieved.
 The conditional optimizer is shown as a red point and the best estimation of the optimizer is shown as a blue point.
@@
 This allows: (i) accounting for the information on each constraint separately and (ii) solving the problems, in which the function values may be undefined out of the feasible domain.
-It should be noted, that the index method can be efficiently parallelized for accelerators \cite{Barkalov2015,Barkalov2016}.
+%It should be noted, that the index method can be efficiently parallelized for accelerators \cite{Barkalov2015,Barkalov2016}.
 

### 2018-04-24 2018/Journal of Supercomputing/JoS_Paper.tex

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{multirow}
%
%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Parallel Algorithm for Solving Constrained Global Optimization Problems
\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Konstantin Barkalov          \and
        Ilya Lebedev %etc.
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{K. Barkalov  \at
               Institute of Information Technology, Mathematics and Mechanics\\ Lobachevsky State University of Nizhni Novgorod\\ Nizhni Novgorod, Russia \\
              \email{konstantin.barkalov@itmm.unn.ru} 
           \and
           I. Lebedev \at
					Institute of Information Technology, Mathematics and Mechanics\\ Lobachevsky State University of Nizhni Novgorod\\ Nizhni Novgorod, Russia \\
							\email{ilya.lebedev@itmm.unn.ru} 
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
This work considers a parallel algorithm for solving multiextremal problems with non-convex constraints. The distinctive feature of this algorithm, which does not use penalty functions, is the separate consideration of each problem constraint. The search process can be conducted by reducing the original multidimensional problem to a number of related one-dimensional problems and solving this set of problems in parallel. An experimental assessment of parallel algorithm efficiency was conducted by finding the numeric solution to several hundred randomly generated multidimensional multiextremal problems with non-convex constraints.
\keywords{Global optimization \and Constrained problems \and Non-convex constraints \and Dimensionality reduction \and Parallel algorithms}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}

This work considers parallel methods for solving global optimization problems with non-convex constraints. The objective function and constraints are assumed to satisfy the Lipschitz condition with a priori unknown Lipschitz constants. The analytical form of the problem’s functions may also be unknown, i.e. they can be set by an algorithm computing their values at various points within the search domain (so-called ``black-box'' functions). Moreover, it is supposed that even a single computation of the function value can be time-consuming, as in the applied problems it requires performing numerical simulation. These assumptions are typical for many approaches to building parallel algorithms for unconstrained global optimization \cite{Jones2001,Paulavicius2011,Evtushenko2013}.

At the same time, it is common for applied constrained optimization problems to be in a situation where violating one constraint results in all other functions returning indeterminate values. An example includes optimal control problems, described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2011,Balandin2017}. It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e. every eigenvalue of $A$ has strictly negative real part. Otherwise the value of the criteria is indeterminate.

This partial computability of the functions in constrained optimization problems substantially complicates the application of the well-known penalty function method (in some cases making it completely impossible). Thanks to its simplicity, this method is one of the most popular approaches to solving problems with constraints. However, calculating the penalty function requires first calculating the values of all of the problem’s functions at the given point, which is impossible if they are partially indeterminate.

In this work, the authors consider an approach to minimizing multiextremal functions under non-convex constraints, developed in \cite{Strongin2000,Sergeyev2001,Barkalov2002} and called \textit{the index method}. The approach is based on separate considering every constraint in the problem and is not related to using penalty functions. According to the index method, each iteration (\textit{a trial}) at a respective point in the search domain includes a sequential check of the problem constraints at that point. As soon as the first constraint violation is found, the trial is interrupted and the method proceeds with the next iteration; no other problem functions are calculated at that point. This allows problems to be solved in which function values may not be determined for the entire search domain. Under this approach, solving multidimensional problems is reduced (using Peano-type space-filling curves) to solving equivalent one-dimensional problems.

It should be noted that standard approaches to algorithm parallelization are not quite applicable to global optimization. For example, the rules for selecting another iteration point are quite simple and do not require parallelization (as overheads associated with organizing parallel computations will nullify any possible acceleration). Some acceleration can be achieved by parallelizing the computation of function values describing the object to be optimized; however, this approach is specific to each individual problem being solved.

The following approach looks more promising. The algorithm can be modified to run several trials in parallel. This approach provides the efficiency (as parallelization is applied to the most computation-intensive part of the problem solving process) and generality (in that it applies to a wide range of global optimization algorithms). The approach, described in \cite{Sidorov2015} for unconstrained optimization, was used in this work for parallelizing constrained optimization algorithms.

The main part of the paper has the following structure. Section 2 states the constrained optimization problem, reviews the index method and an approach to reducing dimensionality by using Peano curves. Section 3 presents a parallel implementation of the index method using a set of space-filling curves. Section 4 presents the results of numerical experiments. Section 5 concludes the paper.

\section{Problem Statement}
\label{sec:2}

Let us consider the $N$-dimensional global optimization problem
\begin{equation}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
\end{equation}
\begin{equation}\label{D}
D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.
\end{equation}
The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy Lipschitz condition 
\[
\left|g_i(y_1)-g_i(y_2)\right|\leq L_i\left\|y_1-y_2\right\|, \;1\leq i\leq m+1, \; y_1,y_2 \in D,\;
\]
with a priory unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal. It is admitted that functions $g_i(y)$ are defined and computable only at the points $y \in D$ satisfying the conditions
\begin{equation}\label{g_k}
g_k(y) \leq 0, \; 1 \leq k < i.
\end{equation}

Employing the continuous single-valued Peano curve $y(x)$ mapping the unit interval $[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}) it is possible to find the minimum in (\ref{problem}) by solving the one-dimensional problem
\[
\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
\]
Algorithms for numerical construction of Peano curve approximation (\textit{evolvent}) are given in \cite{Sergeyev2013}. Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined and computable in the domains 
\[
Q_1=[0,1], \; Q_{i+1}=\left\{x \in Q_i : g_i(y(x)) \leq 0 \right\}, \; 1 \leq i \leq m.
\]

These conditions allows to introduce a classification of the points $x \in [0,1]$ according to the number $\nu (x)$ of the constraints computed at this point. The \textit{index} $\nu(x)$ can also be defined by the conditions
\[
g_i(y(x)) \leq 0, \; 1 \leq i < \nu, \; g_\nu(y(x))>0,
\]
where the last inequality is inessential if $\nu=m+1$.

The considered dimensionality reduction scheme juxtaposes to a multidimensional problem with lipschitzian functions a one-dimensional problem, where the corresponding functions satisfy uniform H{\"o}lder condition (see \cite{Sergeyev2013}), i.e.,
\[
\left|g_i(y(x_1))-g_i (y(x_2))\right| \leq H_i \left|x_1-x_2 \right|^{1/N}, \; x_1,x_2\in [0,1], \; 1\leq i \leq m+1.
\]
Here $N$ is the dimensionality of the initial multidimensional problem and the coefficients $H_i$ are related with Lipschitz constant $L_i$ of the initial problem as $H_i \leq 2L_i \sqrt{N+3}$.

Thus, \textit{a trial} at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm will consist in the following sequence of operations:
\begin{itemize}
	\item Determine the \textit{image} $y^k=y(x^k)$ in accordance with the mapping $y(x)$;
	\item Compute the values $g_1(y^k),...,g_\nu(y^k),$ where the index $\nu \leq m$ is determined by the conditions
	\[
		g_i(y^k )\leq 0, \; 1 \leq i < \nu, \; g_\nu(y^k)>0, \; \nu \leq m.
	\]
\end{itemize}
The occurrence of the first violation of the constraint terminates the trial. In the case, when the point $y^k$ is a feasible one, i.e. when $x^k \in Q_{m+1},$ the trial includes the computation of the values of all problem functions and the index is accepted to be $\nu=m+1$. The pair of values
\[
\nu=\nu(x^k), \; z^k=g_\nu(y(x^k))
\]
is a \textit{trial result}.

The scheme of the serial index algorithm is as follows. The first trial is executed at an arbitrary internal point $x_1 \in (0,1)$. The selection of the point $x^{k+1}, \; k \geq 1,$ of any next trial is determined by the following steps.

Step 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the lower indices in increasing order of the coordinate values, i.e.
\[
0=x_0<x_1<\dots <x_k<x_{k+1}=1,
\]
and juxtapose to them the values $z_i=g_\nu(y(x_i)), \; \nu=\nu(x_i), \; 1 \leq i \leq k,$ computed at these points. The points $x_0=0$ and $x_{k+1}=1$ are introduced additionally and the values $z_0$ and $z_{k+1}$ are not defined.

Rule 2. For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} $R(i)$ using some formulae.

Rule 3. Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
\begin{equation}\label{MaxR}
R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
\end{equation}

Rule 4. Execute the next trial in the internal point of the interval $(x_{t-1},x_t)$, i.e. $x^{k+1} \in (x_{t-1},x_t)$.

Step 5. Check termination condition $\left|x_t-x_{t-1}\right|^{1/N}\leq \epsilon$, where $t$ is the number of interval with the maximum characteristic and $\epsilon > 0$ is the predefined accuracy.

Detailed description of this algorithm and the corresponding theory of convergence are presented in \cite{Strongin2000,Sergeyev2001,Barkalov2002}.

\section{Parallel Index Algorithm with the Set of Evolvents}
\label{sec:3}

The reduction of the multidimensional problems to the one-dimensional ones using evolvents has such important properties as the continuity and preservation of boundedness of function divided differences. However, a partial loss of information on the nearness of the points in the multidimensional space takes place since a point $x \in [0,1]$ has only the left and the right neighbors while the corresponding point $y(x) \in R^N$ has the neighbors in $2N$ directions. As a result, when using the mappings like Peano curve the images $y', y''$, which are close to each other in the $N$-dimensional space can correspond to the preimages $x', x''$, which can be far away from each other in the interval $[0,1]$. This property results in the excess computations since several limit points $x', x''$ of the trial sequence generated by the index method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional space.

One of the possible ways to overcome this disadvantage consists in using the multiple mapping $Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of single evolvent $y(x)$. To construct the set $Y^S(x)$ different approaches can be used. For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^i(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$. The set of Peano curves thus constructed allows one to obtain $y', y''$ from $D$ for any close multidimensional images, which differ only in one coordinate, close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$.

Using the multiple mapping allows solving initial problem (\ref{problem}) by parallel solving the problems
\[
\min \left\{\varphi(y^s(x)): x \in [0,1], \; g_i(y^s(x))\leq 0, \; 1 \leq i \leq m\right\},\; 1\leq s \leq S.
\]
on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a separate processor/node. The trial results at the point $x^k$ obtained for the problem being solved by particular processor are interpreted as the results of the trials in the rest problems (at the corresponding points $x^{k_1},...,x^{k_S}$). In this approach, a trial at the point $x^k \in [0,1]$ executed in the framework of the $s$-th problem, consists in the following sequence of operations.

\begin{enumerate}

	\item Determine the image $y^k=y^s(x^k)$ for the evolvent $y^s(x)$.

	\item Inform the rest of processors about the start of the trial execution at the point $y^k$ (\textit{the blocking} of the point $y^k$).

	\item Compute the values $g_1(y^k),...,g_\nu(y^k)$, where the index $\nu \leq m$ is determined by the conditions
	\[
		g_i(y^k)\leq 0, \; 1 \leq i < \nu, \; g_\nu(y^k)>0, \; \nu \leq m.
	\]
The occurrence of the first violation of the constraint terminates the trial. In the case when the point $y^k$ is a feasible one the trial includes the computation of the values of all problem functions. In this situation, the index is set to $\nu=m+1$. The triplet
\[
y^s(x^k), \; \nu=\nu(x^k), \; z^k=g_\nu(y^s(x^k))
\]
is the result of the trial at the point $x^k$.

\item Determine the preimages $x^{k_s} \in [0,1]$, $1 \leq s \leq S$, of the point $y^k$ and interpret the trial executed at the point $y^k \in D$ as the execution of the trials in the $S$ points  $x^{k_1},...,x^{k_S}$ with the same results
\[
 \nu(x^{k_1}) = ... = \nu(x^{k_S}) = \nu(x^k),
\]
\[
 g_\nu(y^1(x^{k_1})) = ... = g_\nu(y^S(x^{k_S})) = z^k.
\]

\item Inform the rest of processors about the trial results at the point $y^k$.

\end{enumerate}

The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the sequential algorithm (except the method of the trial execution). For the organization of the interactions among the processors, the queues are created on each processor, where the processors store the information on the executed iterations in the form of the tuples: the processor number $s$, the trial point $x^{k_s}$, the index $\nu(x^{k_s})$, and the value $g_\nu(y^s (x^{k_s} ) )$. Moreover, the index of the blocked point is assumed to be equal to $-1$; the function value at this point is undefined.

The proposed parallelization scheme was implemented with the use of MPI technology. Main features of implementation consist in the following. A separate MPI-process is created for each of $S$ one-dimensional problems being solved, usually, one process per one processor employed. Each process can use $p$ threads, usually one thread per an accessible core.

At every iteration of the method, the process with the index $s, 0\leq s < S,$ performs $p$ trials in parallel at the points $x^{s+iS}, 0\leq i < p$. At that, each process stores all $S\cdot p$ points, and an attribute indicating whether this point is blocked by another process or not is stored for each point. Let us remind that the point is blocked if the process starts the execution of a trial at this point.

At every iteration of the algorithm, operating within the $s$-th process, determines the coordinates of $p$ ``its own'' trial points. Then, the interchange of the coordinates of images of the trial points $y^{s+iS},\; 0\leq i<p,\; 0\leq s< S$ is performed (from each process to each one). After that, the preimages $x^{q+iS},\; 0\leq q<S,\; q\neq s$ of the points received by the $s$-th process from the neighbor ones are determined with the use of the evolvent $y^s(x)$. The points blocked within the $s$-th process will correspond to the preimages obtained. Then, each process performs the trials at the non-blocked points, the computations are performed in parallel using OpenMP. The results of the executed trials (the index of the point, the computed values of the problem functions, and the attribute of unblocking of this point) are transferred to all rest processes. All the points are added to the search information database, and the transition to the next iteration is performed.

\section{Results of Numerical Experiments}
\label{sec:4}

A well-known approach to the investigation and comparing of the multiextremal optimization algorithms is based on testing these methods by solving a set of problems, chosen randomly from some specially designed class.

GKLS generator for the functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.) has been proposed in \cite{Gaviano2003}. Four GKLS classes of differentiable test functions of the dimensions $N = 4$ and $5$, have been used. For each dimension, both \textit{Hard} and \textit{Simple} classes have been considered. The difficulty of a class was increased either by decreasing the radius of the attraction region of the global minimizer, or by decreasing the distance from the global minimizer $y^*$ to the domain boundaries. Application of the generator for studying some optimization algorithms has been described in \cite{Paulavicius2014,Sergeyev2015}.

In this study we will use GKLS generator to produce the constrained problems. The scheme that allows to form the constrained global optimization problems is proposed in \cite{Gergel2017}. In the previous investigations, the index method has been confirmed experimentally to be not inferior to well-known analogues. The comparing of the method to well known DIRECT one \cite{Jones2001} in solving the unconstrained optimization problems has been performed in \cite{Barkalov2015,Barkalov2016}. In the present study, an experimental investigation of the speedup, which is obtained by the use of the index method in combination with the two-level parallelization scheme from \cite{Sidorov2015}.

The experiments have been carried out by solving a series of $100$ problems with two constraints and the objective functions from the \textit{Simple} and \textit{Hard} GKLS classes with the dimensionalities $N=4$, $N=5$. The number of the used cluster nodes $S$ and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at each node have been varied. The problem was considered to be solved, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^*\right\|\leq \delta$. The size of the vicinity was selected as $\delta=0.03\left\|b-a\right\|$, where $a$ and $b$ are borders of the search domain. For the purpose of simulation of the computational complexity inherent to applied problems of optimization, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (series summation of $80$ thousand elements).

The average time and number of iterations, which were required to solve the problems of the series at various parallelization parameters are reflected in Tables 1 and 2. Here \textit{Node(core)} are the numbers of employed nodes and cores per a node, correspondingly.

\begin{table}
	\caption{Average time}
	\label{tab:1}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
		\multirow{2}{*}{\textit{Node(core)}} & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
1(1)	&	220.5	&	334.8	&	&	1223.6	&	1386.6	\\
1(16)	&	31.3	&	49.1	&	&	211.8	&	547.2	\\
2(1)	&	158.4	&	260	&	&	1052.9	&	1458.1	\\
2(16)	&	22.1	&	35.9	&	&	227.5	&	603	\\
4(1)	&	127.7	&	286.4	&	&	951.3	&	1362.2	\\
4(16)	&	20.9	&	45	&	&	206	&	925.7	\\
8(1)	&	99.3	&	141.8	&	&	700.1	&	897.3	\\
8(16)	&	31	&	77.7	&	&	264.6	&	374	\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Average number of iterations}
	\label{tab:2}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
		\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
1(1)	&	58320	&	84546	&	&	266943	&	287102	\\
1(16)	&	4297	&	6601	&	&	22655	&	56754	\\
2(1)	&	34791	&	52126	&	&	188465	&	241369	\\
2(16)	&	2029	&	3239	&	&	16689	&	40763	\\
4(1)	&	22223	&	47771	&	&	135734	&	180489	\\
4(16)	&	1281	&	2483	&	&	9241	&	35024	\\
8(1)	&	13844	&	18933	&	&	77748	&	94563	\\
8(16)	&	608	&	1473	&	&	5820	&	23033	\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

The results demonstrate the presence of the speedup when using shared memory at a node (performing several trials within a problem in parallel) as well as the distributed memory (parallel solving of several subproblems at different nodes). At that, the highest time speedup was 10 (when using 64 cored on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster nodes). The difference in the speedups in time and in the number of iterations can be explained by the effect of the overheads of the data transmission between the processes. Note that when solving the applied optimization problems, the computing of the problem function values even in one point is a computation costly operation. The data transfer overheads will not affect the total computational costs predominately in this case, and the time speedup will not differ from the iteration one so strongly.

\section{Conclusions}
\label{sec:5}

The parallel index method for solving constrained global optimization problems considered in the present work:
\begin{itemize}
	\item allows solving the initial problem directly, without the use of the penalty functions (thus, the issues of selection the penalty coefficient and of solving a series of unconstrained problems with different penalty coefficients are eliminated);
	\item allows solving the problems, which the values of the problem function are not defined everywhere (for example, the objective function values are undefined out of the feasible domains of the problem constraints);
	\item allows using the two-level parallelization scheme with the shared and distributed memory proposed earlier for the unconstrained optimization methods.
\end{itemize}

The parallel algorithm has demonstrated speedup with respect to the number of processors/cores employed. This was confirmed by the results of the numerical solving of several hundred test problems using 128 cores of Lobachevsky supercomputer. 
%The direction of further research is the generalization of the considered parallel algorithm for solving multicriteria problems.

% Non-BibTeX users please use
\begin{thebibliography}{}

\bibitem{Jones2001}
Jones, D.R.: The direct global optimization algorithm. In: Floudas, C.A., Pardalos, P.M. (eds.) The Encyclopedia of Optimization, Second Edition. pp. 725--735. Springer (2009)

%\bibitem{Evtushenko2009} Evtushenko, Yu.G., Malkova, V.U., Stanevichyus, A.A.: Parallel global optimization of functions of several variables. Comput. Math. Math. Phys. 49(2), 246--260 (2009)

\bibitem{Paulavicius2011}
Paulavi\v cius, R., \v Zilinskas, J., Grothey, A.: Parallel branch and bound for global optimization with combination of Lipschitz bounds. Optim. Methods Softw. 26(3), 487--498 (2011)

\bibitem{Evtushenko2013}
Evtushenko, Y.G., Posypkin, M.A.: A deterministic approach to global box-constrained optimization. Optim. Lett. 7(4), 819--829 (2013)

\bibitem{Balandin2011}
Balandin, D.V., Kogan, M.M.: Optimal linear-quadratic control: From matrix equations to linear matrix inequalities. Automation and Remote Control. 72(11), 2276--2284 (2011)

\bibitem{Balandin2017}
Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 (2017)

\bibitem{Strongin2000}
Strongin, R.G., Sergeyev, Ya.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000)

\bibitem{Sergeyev2001}
Sergeyev, Y.D., Famularo D., Pugliese P. Index Branch-and-Bound Algorithm for Lipschitz univariate global optimization with multiextremal constraints. J. Glob. Optim. 21(3), 317--341 (2001)

\bibitem{Barkalov2002}
Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)

\bibitem{Sidorov2015}
Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of computationally intensive multiextremal optimization problems. Lecture Notes in Computer Science. 9251, 505--515 (2015)

\bibitem{Sergeyev2013}
Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting Space-Filling Curves. Springer (2013)

\bibitem{Gaviano2003}
Gaviano, M., Lera, D., Kvasov, D.E., Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Trans. Math. Software 29, 469--480 (2003)

%\bibitem{Sergeyev2006} Sergeyev, Ya.D., Kvasov, D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim. 16(3), 910--937 (2006)

\bibitem{Paulavicius2014}
Paulavi\v cius, R., Sergeyev, Y.D., Kvasov, D.E., \v Zilinskas, J. Globally-biased \textsc{Disimpl} algorithm for expensive global optimization. J. Glob. Optim. 59(2), 545--567 (2014)

\bibitem{Sergeyev2015}
Sergeyev, Y.D., Kvasov, D.E.: A deterministic global optimization using smooth diagonal auxiliary functions. Communications in Nonlinear Science and Numerical Simulation. 21(1-3), 99–111 (2015)

\bibitem{Gergel2017}
Gergel, V.: An approach for generating test problems of constrained global optimization. Lecture Notes in Computer Science. 10556, 314--319 (2017)

\bibitem{Barkalov2015}
Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi coprocessor for solving global optimization problems. Lecture Notes in Computer Science. 9251, 307--318 (2015)

\bibitem{Barkalov2016}
Barkalov, K., Gergel, V.: Parallel global optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016)

%
%\bibitem{RefJ}
%% Format for Journal Reference
%Author, Article title, Journal, Volume, page numbers (year)
%% Format for books
%\bibitem{RefB}
%Author, Book title, page numbers. Publisher, place (year)
% etc
\end{thebibliography}

\end{document}
% end of file template.tex

### 2018-04-24 2018/LeGo/LeGo.tex
enumeration is corrected
Last russian version

@@
 \usepackage{graphicx}
-\usepackage{enumerate}
-\pdfmapfile{+txfonts.map}
+\usepackage{enumitem}
+%\pdfmapfile{+txfonts.map}
 
@@
 % Title portion
-\title{The Title Goes Here with Each Initial Letter Capitalized}
 
-\author{Victor Gergel\corref{cor1}}
-\author{Konstantin Barkalov}
+\title{A Flexible  Generator of Constrained Global Optimization Test Problems}
+
+\author{Victor Gergel}
+\author{Konstantin Barkalov\corref{cor1}}
 %\eaddress[url]{http://www.aip.org}
@@
 \author{Maria Rachinskaya}
+\author{Alexander Sysoyev}
 
 \affil{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia.}
-\corresp[cor1]{Corresponding author: gergel@unn.ru}
+
+\corresp[cor1]{Corresponding author: konstantin.barkalov@itmm.unn.ru}
 
@@
 \begin{abstract}
+\Russian
 В работе рассматривается дальнейшее развитие подхода к построению тестовых задач глобальной оптимизации с нелинейными ограничениями.
@@
 The most important drawback consists of the fact that the constraints are absent in the proposed test problems as a rule (or the constraints are relatively simple: linear, convex, etc.).
-
+\Russian
 В данной работе проводится дальнейшее развитие подхода для порождения задач глобальной оптимизации с нелинейными ограничениями, первоначально предложенного в \cite{Gergel2017}.
@@
 \end{itemize}
-Указанные возможности позволяют на более высоком уровне имитировать свойства прикладных задач условной глобальной оптимизации.
 
-\section{Test problem classes}
+Указанные возможности позволяют более качественно имитировать свойства прикладных задач условной глобальной оптимизации.
+
+\section{TEST PROBLEM CLASSES}
 A well-known approach to investigating and comparing the multiextremal optimization algorithms is based on testing these methods by solving a set of problems, chosen randomly from some specially designed class.
 
-For example, one generator for random samples of two-dimensional test functions has been described in \cite{Grishagin2015}. This generator produces two-dimensional functions according to the formula
-\begin{equation} \label{VAG}
-\varphi(y)= -\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2 + \left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},
-\end{equation}
-where $g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2)$, $h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2)$, $y=(y_1,y_2)\in R^2, \; 0 \leq y_1,y_2 \leq 1$, and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
-
+For example, one generator for random samples of two-dimensional test functions has been described in \cite{Grishagin2015}. 
+%This generator produces two-dimensional functions according to the formula
+%\begin{equation} \label{VAG}
+%\varphi(y)= -\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2 + \left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},
+%\end{equation}
+%where $g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2)$, $h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2)$, $y=(y_1,y_2)\in R^2, \; 0 \leq y_1,y_2 \leq 1$, and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
 Another generator (\textit{GKLS generator}) for the functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.) has been proposed in \cite{Gaviano2003}.
@@
 \Russian
+
 Одним из важных свойств, характеризующих прикладные задачи условной оптимизации, является расположение условного минимума на границе feasible domain.
 При генерации задач по схеме, описанной в \cite{Gergel2017}, условный минимум будет, вообще говоря, находиться внутри feasible domain.
-Поэтому в новой версиии генератора задач была добавлена возможность как смещения условного минимума на границу, так и выбора числа активных ограничений в нем.
+Поэтому в новой версиии генератора задач была добавлена возможность как смещения условного минимума на границу, так и выбора числа active constraints в нем.
 
@@
 
-В случае, если нужно сгенерировать задачу с заданным числом $S$ активных ограничений в точке условного минимума, то тогда правила меняются следующим образом 
+В случае, если нужно сгенерировать задачу с заданным числом $S$ активных ограничений в точке условного минимума, то тогда правила меняются следующим образом.
 \begin{enumerate}
@@
 
-\section{Numerical results}
+\section{NUMERICAL RESULTS}
 
@@
 The feasible domains are highlighted by green.
+\Russian
+Fig.~\ref{fig:VAG} (a,b,c,d) соответствуют разным способам генерации тестовых задач.
 
-Fig.~\ref{fig:VAG} (a,b,c,d) соответствуют разным способам генерации тестовых задач.
-\begin{description}
-	\item[(a)] Задача сгенерирована с расположением constrained optimizer внутри feasible domain.
-	\item[(b)] Модификация задачи из п. (a): unconstrained optimizer находится вне feasible domain.
-	\item[(c)] Модификация задачи из п. (b): constrained optimizer находится на границе feasible domain, число активных ограничений не задано.
-	\item[(d)] Модификация задачи из п. (c): число активных ограничений задано равным двум.
-\end{description}
+\begin{enumerate}[label=(\alph*)]
+	\item Задача сгенерирована с расположением constrained optimizer внутри feasible domain.
+	\item Модификация задачи из п. (a): unconstrained optimizer находится вне feasible domain.
+	\item Модификация задачи из п. (b): constrained optimizer находится на границе feasible domain, число активных ограничений не задано.
+	\item Модификация задачи из п. (c): число активных ограничений задано равным двум.
+\end{enumerate}
 
-Fig.~\ref{fig:VAG} (a,b,c,d) also shows the points of 628 and 764 trials, correspondingly, performed by the \textit{index method} for solving constrained global optimization problems until the required accuracy $\epsilon=10^{-2}$ was achieved.
+Fig.~\ref{fig:VAG} (a,b,c,d) also shows the points of XXX, YYY, ZZZ and TTT trials, correspondingly, performed by the \textit{index method} for solving constrained global optimization problems until the required accuracy $\epsilon=10^{-2}$ was achieved.
 The conditional optimizer is shown as a red point and the best estimation of the optimizer is shown as a blue point.
@@
 \section {Conclusion}
-
+\Russian
 This paper considers the method for generating global optimization test problems with non-convex constraints that allows:
@@
 	\item to known a priori the conditional global minimizer of the objective function;
-	\item to generate the unconditional global minimizer of the objective function out of the feasible domain (to simulate the constraints and objective function in the applied optimization problems).
+
+	\item to generate the unconditional global minimizer of the objective function out of the feasible domain;
+	\item генерировать задачи с точкой условного минимума, расположенной на границе допустимой области;
+	\item задавать число ограничений, активных в точке оптимума.
 \end{itemize}

### 2018-04-27 2018/LeGo/LeGo.tex
Translate to English and correct figures

@@
 \usepackage{rotating}
+\usepackage{subfig}
 \usepackage{graphicx}
+\usepackage{tabularx}
 \usepackage{enumitem}
+
 %\pdfmapfile{+txfonts.map}
 
-%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
-\usepackage[utf8x]{inputenc}
-\usepackage[english,russian]{babel}
-\usepackage{cmap}
+%%%%%ÄÎÁÀÂÈË ÄËß ÐÓÑÑÊÎÃÎ ÒÅÊÑÒÀ
+%\usepackage[utf8x]{inputenc}
+%\usepackage[english,russian]{babel}
+%\usepackage{cmap}
 %%%%%
@@
 
-\affil{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia.}
+\affil{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia}
 
@@
 \begin{abstract}
-\Russian
-В работе рассматривается дальнейшее развитие подхода к построению тестовых задач глобальной оптимизации с нелинейными ограничениями.
-Для порождаемых задач расположение глобального минимума является известным.
-Отличительной чертой рассматриваемого генератора является возможность указания желаемого количества ограничений и доли допустимой области относительно все области глобального поиска.
-В обновленной версии генератора добавлена возможность формирования задач с условным минимумом, расположенным на границе допустимой области, и указания числа ограничений, активных в точке минимума.
-Демонстрация разработанного подхода приводится на примере известного индексного метода для решения сложных многоэкстремальных задач оптимизации с нелинейными ограничениями.
+In the present work, further development of an approach to constructing the test global 
+optimization problems with nonlinear constraints is considered. In the generated problems, the 
+location of the global minimum is known. The considered generator is featured by an option of 
+specifying desired number of constraints and the fraction of feasible domain relative to the 
+whole global search domain. In the updated version of the generator, the options of forming the 
+problems with a conditional minimum located at the boundary of the feasible domain and of 
+specifying the number of constraints active in the minimum points have been added. The 
+demonstration of the developed approach is given on the example of a known index method for 
+solving the complex multiextremal optimization problems with nonlinear constraints.
 \end{abstract}
@@
 
-In the present paper, the methods for generating the global optimization test problems with non-convex constraints
+In the present paper, the methods for generating the global optimization test problems with non-
+convex constraints
 \begin{eqnarray}
-&\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}, \label{i_problem} \\
+&\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}, 
+\label{i_problem} \\
 &D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\} \label{D}
 \end{eqnarray}
-are considered.
-The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints are supposed to satisfy the Lipschitz condition
+are considered. The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and 
+the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints are supposed to satisfy the 
+Lipschitz condition
 \[ \left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq i \leq m+1. \]
-with the Lipschitz constants unknown a priori.
-The analytical formulae of the problem functions may be unknown, i.e.
-these ones may be defined by an algorithm for computing the function values in the search domain (so called ``black-box''-functions).
-It is supposed that even a single computing of a problem function value may be a time-consuming operation since it is related to the necessity of numerical modeling in the applied problems (see, for example, \cite{Menniti2008,Kvasov2015,Modorskii2017}).
+with the Lipschitz constants unknown a priori. The analytical formulae of the problem functions 
+may be unknown, i.e. these ones may be defined by an algorithm for computing the function 
+values in the search domain (so called ``black-box''-functions). It is supposed that even a single 
+computing of a problem function value may be a time-consuming operation since it is related to 
+the necessity of numerical modeling in the applied problems (see, for example, 
+\cite{Menniti2008,Kvasov2015,Modorskii2017}).
 
-The evaluation of efficiency of the developed methods is one of the key problems in the optimization theory and applications.
-Unfortunately, it is difficult to obtain any theoretical estimates in many cases.
-As a result, the comparison of the methods is performed by carrying out the computational experiments on solving some test optimization problems in most cases.
-In order to obtain a reliable evaluation of the efficiency of the methods, the sets of test problems should be diverse and representative enough.
-The problem of choice of the test problems has been considered in a lot of works (see, for example, \cite{Floudas1999,Gaviano2003,Ali2005,Addis2007}).
-Unfortunately, in many cases, the proposed sets contain a small number of test problems, and it is difficult to obtain the problems with desired properties.
-The most important drawback consists of the fact that the constraints are absent in the proposed test problems as a rule (or the constraints are relatively simple: linear, convex, etc.).
-\Russian
-В данной работе проводится дальнейшее развитие подхода для порождения задач глобальной оптимизации с нелинейными ограничениями, первоначально предложенного в \cite{Gergel2017}.
-При генерации тестовых задач может быть указано необходимое количество ограничений и желаемая доля допустимой области относительно все области глобального поиска.
-При этом для порождаемых задач расположение глобального минимума является известным, что существенно упрощает оценку результатов выполненных вычислительных экспериментов.
-В очередную версию генератора, в дополнение к указанным свойствам, были добавлены следующие возможности.
+The evaluation of efficiency of the developed methods is one of the key problems in the 
+optimization theory and applications. Unfortunately, it is difficult to obtain any theoretical 
+estimates in many cases. As a result, the comparison of the methods is performed by carrying 
+out the computational experiments on solving some test optimization problems in most cases. In 
+order to obtain a reliable evaluation of the efficiency of the methods, the sets of test problems 
+should be diverse and representative enough. The problem of choice of the test problems has 
+been considered in a lot of works (see, for example, 
+\cite{Floudas1999,Gaviano2003,Ali2005,Addis2007}). Unfortunately, in many cases, the 
+proposed sets contain a small number of test problems, and it is difficult to obtain the problems 
+with desired properties. The most important drawback consists of the fact that the constraints 
+are absent in the proposed test problems as a rule (or the constraints are relatively simple: linear, 
+convex, etc.).
+
+In the present work, further development of the approach to generating the global optimization 
+problems with nonlinear constraints proposed originally in Ref.~\cite{Gergel2017} has been 
+conducted.
+When generating the test problems, the necessary number of constraints and the desired fraction 
+of the feasible domain relative to the whole global search domain can be specified. 
+At that, the location of the global minimum is known for the generated problems that simplifies 
+evaluating the results of the numerical experiments essentially. 
+in addition to the specified options, 
+the following functions have been added in the current version of the generator:
+
 \begin{itemize}
-	\item Возможность генерации задач с точкой условного минимума, расположенной на границе допустимой области.
-	\item Возможность указания числа ограничений, активных в точке оптимума.
+\item The possibility to generate the problems with the conditional minimum point located at the 
+boundary of the feasible domain.
+\item The possibility to specify the number of constraints active in the optimum point.
 \end{itemize}
 
-Указанные возможности позволяют более качественно имитировать свойства прикладных задач условной глобальной оптимизации.
+The above capabilities allow emulating the properties of the applied constrained global 
+optimization problems more adequately.
 
 \section{TEST PROBLEM CLASSES}
-A well-known approach to investigating and comparing the multiextremal optimization algorithms is based on testing these methods by solving a set of problems, chosen randomly from some specially designed class.
+A well-known approach to investigating and comparing the multiextremal optimization 
+algorithms is based on testing these methods by solving a set of problems, chosen randomly 
+from some specially designed class.
 
-For example, one generator for random samples of two-dimensional test functions has been described in \cite{Grishagin2015}. 
+For example, one generator for random samples of two-dimensional test functions has been 
+described in \cite{Grishagin2015}. 
 %This generator produces two-dimensional functions according to the formula
 %\begin{equation} \label{VAG}
-%\varphi(y)= -\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2 + \left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},
+%\varphi(y)= -
+%\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2 + 
+%\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},
 %\end{equation}
-%where $g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2)$, $h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2)$, $y=(y_1,y_2)\in R^2, \; 0 \leq y_1,y_2 \leq 1$, and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
-Another generator (\textit{GKLS generator}) for the functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.) has been proposed in \cite{Gaviano2003}.
-Application of this generator for studying some unconstrained optimization algorithms has been described in \cite{Paulavicius2014,Sergeyev2015}.
+%where $g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2)$, $h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2)$, 
+%$y=(y_1,y_2)\in R^2, \; 0 \leq y_1,y_2 \leq 1$, and coefficients $A_{ij}, B_{ij}, C_{ij}, 
+%D_{ij}$  are taken uniformly in the interval $[-1,1]$.
+Another generator (\textit{GKLS generator}) for the functions of arbitrary dimensionality with 
+known properties (the number of local minima, the size of their domains of attraction, the global 
+minimizer, etc.) has been proposed in \cite{Gaviano2003}.  Application of this generator for 
+studying some unconstrained optimization algorithms has been described in 
+\cite{Paulavicius2014,Sergeyev2015}.
 
-The generator GCGen (Global Constrained optimization problem Generator) which allows to generate the test global optimization problems with $m$ constraints на основе функций заданного класса, has been proposed in \cite{Gergel2017}.
-В данной статье были были предложены the rules, which allow formulating the constrained global optimization problems so that:
+The generator GCGen (Global Constrained optimization problem Generator) which allows to 
+generate the test global optimization problems with $m$ constraints based on the functions of 
+this class has been proposed in Ref.~\cite{Gergel2017}. In the present paper, the rules, which 
+allow formulating the constrained global optimization problems have been proposed so that:
 \begin{itemize}
-	\item one could control the size of feasible domain with respect to the whole domain of the parameters' variation;
-	\item the global minimizer of the objective function would be known a priori taking into account the constraints;
-	\item the global minimizer of the objective function without accounting for the constraints would be out of the feasible domain.
+	\item one could control the size of feasible domain with respect to the whole domain of 
+the parameters' variation;
+	\item the global minimizer of the objective function would be known a priori taking into 
+account the constraints;
+	\item the global minimizer of the objective function without accounting for the 
+constraints would be out of the feasible domain.
 \end{itemize}
 
-\Russian
+The location of the conditional minimum at the boundary of the feasible domain is an important 
+property featuring the applied constrained optimization problems. 
+When generating the problems according to the scheme described in Ref.~\cite{Gergel2017}, 
+the conditional minimum, in general, will be located inside the feasible domain. Therefore, the 
+options of shifting the conditional minimum to the boundary of the feasible domain as well of 
+specifying the number of the active constraints in this one have been added in the new version 
+of the problem generator.
 
-Одним из важных свойств, характеризующих прикладные задачи условной оптимизации, является расположение условного минимума на границе feasible domain.
-При генерации задач по схеме, описанной в \cite{Gergel2017}, условный минимум будет, вообще говоря, находиться внутри feasible domain.
-Поэтому в новой версиии генератора задач была добавлена возможность как смещения условного минимума на границу, так и выбора числа active constraints в нем.
-
-Смещение условного минимума на границу feasible domain осуществляется следующим образом.
+The shifting of the conditional minimum to the boundary of the feasible domain is performed in 
+the following way.
 \begin{enumerate}
-	\item Генерируется задача с произвольным расположением условного минимума $\overline{y}$;
-	\item Из точки $\overline{y}$ с заданным шагом $h$ осуществляется покоординатный поиск допустимой точки $y^*$, для которой хотя бы одно из ограничений является активным, т.е.
-существует такой номер $j, 1 \leq j\leq m$, что $\left|g_j(y^*)\right| \leq \delta$.
-	\item Полученное грубое значение $y^*$ уточняется локальным методом. 
-	\item Выполняется преобразование координат, переводящее точку $\overline{y}$ в точку $y^*$. Тем самым целевая функция достигает своего минимума на границе feasible domain. 
+\item A problem with an arbitrary location of the constrained minimum $\overline{y}$ is 
+generated;
+\item From the point $\overline{y}$, a coordinate search of a feasible point $y^*$, for which at 
+least one constraint is active i.e. there is such index $j, 1 \leq j\leq m$ that $\left|g_j(y^*)\right| 
+\leq \delta$ is performed with given step $h$. 
+\item A rough estimate of $y^*$ obtained is refined by a local method. 
+\item The coordinate transformation transferring the point $\overline{y}$ to the point $y^*$ is 
+performed. This way, the objective function reaches its minimum at the boundary of feasible 
+domain. 
 \end{enumerate}
 
-Отметим, что так как поиск точки $y^*$ осуществляется по каждой координате в отдельности, то трудоемкость поиска будет возрастать линейно при увеличении размерности задачи.
+Note that since the search of the point $y^*$ is performed in each coordinate separately, the 
+computation costs would increase linearly with increasing the dimensionality of the problem.
 
-В случае, если нужно сгенерировать задачу с заданным числом $S$ активных ограничений в точке условного минимума, то тогда правила меняются следующим образом.
+In the case, when it is necessary to generate a problem with a predefined number of active 
+constraints $S$ in the constrained minimum point, the rules are modified as follows.
 \begin{enumerate}
-	\item Генерируется задача с произвольным расположением условного минимума $\overline{y}$;
-	\item Из точки $\overline{y}$ с заданным шагом $h$ осуществляется покоординатный поиск допустимой точки $y^*$, для которой хотя бы одно из ограничений является активным, т.е.
-существует такой номер $j, 1 \leq j\leq m$, что $\left|g_j(y^*)\right| \leq \delta$.
-	\item Полученное грубое значение $y^*$ уточняется локальным методом. 
-	\item Определяется число $K$ ограничений, активных в точке $y^*$.
-	\item Если $K>S$, то из числа активных выбираются $K-S$ ограничений, и путем вычитания положительного параметра $q$ из правой части они переводятся в неактивные.
-	\item Если $K<S$, то из числа неактивных выбираются $S-K$ ограничений, и путем прибавления положительного параметра $q$ в правой части они переводятся в активные.
-	\item Выполняется преобразование координат, переводящее точку $\overline{y}$ в точку $y^*$.
-Тем самым целевая функция достигает своего минимума на границе feasible domain с заданным числом активных ограничений.
+\item A problem with an arbitrary location of a conditional minimum $\overline{y}$ is 
+generated;
+\item From the point $\overline{y}$, a coordinate search of a feasible point $y^*$, for which at 
+least one constraint is active i. e. there is such index $j, 1 \leq j\leq m$ that $\left|g_j(y^*)\right| 
+\leq \delta$ is performed with a predefined step $h$.
+\item A rough estimate $y^*$ obtained is refined by a local method. 
+\item The number $K$ of the constraints active at the point $y^*$ is determined.
+\item If $K>S$, $K-S$ constraints are selected from the set of active constraints, and these ones 
+are set to be inactive by subtracting a positive parameter $q$ from the right hand side.
+\item If $K<S$, $S-K$ constraints are selected from the set of the inactive ones, and and these 
+ones are set to be the active by adding a positive parameter $q$ to the right hand side.
+\item The coordinate transformation transferring the point $\overline{y}$ to the point $y^*$ is 
+performed. This way, the objective function reaches its minimum at the boundary of the feasible 
+domain with the predefined number of active constraints. 
 \end{enumerate}
@@
 
-As an illustration, the level lines of the objective functions and the zero-level lines of four constraints for the problem constructed on the base of functions (\ref{VAG}) with the volume fraction of the feasible domains $\Delta = 0.5$ are shown in Fig.~\ref{fig:VAG}.
-The feasible domains are highlighted by green.
-\Russian
-Fig.~\ref{fig:VAG} (a,b,c,d) соответствуют разным способам генерации тестовых задач.
+As an illustration, the level lines of the objective functions and the zero-level lines of four 
+constraints for the problem constructed on the base of functions from \cite{Grishagin2015} with the volume 
+fraction of the feasible domains $\Delta = 0.5$ are shown in Fig.~\ref{fig:VAG}. The feasible 
+domains are highlighted by green.
+
+Fig.~\ref{fig:VAG} (a,b,c,d) correspond to different methods of generating the test problems.
 
 \begin{enumerate}[label=(\alph*)]
-	\item Задача сгенерирована с расположением constrained optimizer внутри feasible domain.
-	\item Модификация задачи из п. (a): unconstrained optimizer находится вне feasible domain.
-	\item Модификация задачи из п. (b): constrained optimizer находится на границе feasible domain, число активных ограничений не задано.
-	\item Модификация задачи из п. (c): число активных ограничений задано равным двум.
+	\item The problem has been generated with the constrained optimizer located inside the 
+feasible domain.
+\item A modification of the problem from item (a): unconstrained optimizer is located outside 
+the feasible domain.
+\item A modification of the problem from item (b): constrained optimizer is located at the 
+boundary of the feasible domain, the number of active constraints is not defined.
+\item A modification of the problem from item (c): the number of active constraints is defined 
+to be two.
 \end{enumerate}
 
-Fig.~\ref{fig:VAG} (a,b,c,d) also shows the points of XXX, YYY, ZZZ and TTT trials, correspondingly, performed by the \textit{index method} for solving constrained global optimization problems until the required accuracy $\epsilon=10^{-2}$ was achieved.
-The conditional optimizer is shown as a red point and the best estimation of the optimizer is shown as a blue point.
-
-The index method has been proposed and developed in \cite{Strongin2000,Sergeyev2001,Barkalov2002}.
-The approach is based on a separate accounting for each constraint of the problem and is not related to the use of the penalty functions.
-According to the rules of index method, every iteration includes a sequential checking of fulfillment of the problem constraints at this point.
-The first occurrence of violation of any constraint terminates the trial and initiates the transition to the next iteration.
-This allows: (i) accounting for the information on each constraint separately and (ii) solving the problems, in which the function values may be undefined out of the feasible domain.
-%It should be noted, that the index method can be efficiently parallelized for accelerators \cite{Barkalov2015,Barkalov2016}.
-
-\begin{figure}
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{vag_04.png} \\ (a)}
-\end{minipage}
-\hfill
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{vag_06.png} \\ (b)}
-\end{minipage}
-\caption{The problems based on functions (\ref{VAG}) }
+\begin{figure}[ht]
+  \centering
+  \begin{tabular}[b]{cc}
+    \includegraphics[width=.44\linewidth]{1.png} & \includegraphics[width=.44\linewidth]{2.png}\\
+    \small (a) & \small (b) \\
+    \includegraphics[width=.44\linewidth]{3.png} & \includegraphics[width=.44\linewidth]{4.png}\\
+    \small (c) & \small (d) 
+  \end{tabular}
+  \caption{The test problems generated by GCGen}
 	\label{fig:VAG}
@@
 
+Fig.~\ref{fig:VAG} (a,b,c,d) also shows the points of 425, 381, 289 and 188 trials, 
+correspondingly, performed by the \textit{index method} for solving constrained global 
+optimization problems until the required accuracy $\epsilon=10^{-2}$ was achieved. The 
+conditional optimizer is shown as a red point and the best estimation of the optimizer is shown 
+as a blue point.
+
+The index method has been proposed and developed in 
+\cite{Strongin2000,Sergeyev2001,Barkalov2002}. The approach is based on a separate 
+accounting for each constraint of the problem and is not related to the use of the penalty 
+functions. According to the rules of index method, every iteration includes a sequential 
+checking of fulfillment of the problem constraints at this point. The first occurrence of violation 
+of any constraint terminates the trial and initiates the transition to the next iteration. This allows: 
+(i) accounting for the information on each constraint separately and (ii) solving the problems, in 
+which the function values may be undefined out of the feasible domain. 
+%It should be noted, that the index method can be efficiently parallelized for accelerators 
+%\cite{Barkalov2015,Barkalov2016}.
+
 \section {Conclusion}
-\Russian
-This paper considers the method for generating global optimization test problems with non-convex constraints that allows:
+
+This paper considers the method for generating the global optimization test problems with non-
+convex constraints, which allows:
 \begin{itemize}
-	\item to control the size of feasible domain with respect to the whole domain of the parameters' variation;
-	\item to known a priori the conditional global minimizer of the objective function;
-
-	\item to generate the unconditional global minimizer of the objective function out of the feasible domain;
-	\item генерировать задачи с точкой условного минимума, расположенной на границе допустимой области;
-	\item задавать число ограничений, активных в точке оптимума.
+	\item controlling the size of the feasible domain with respect to the whole domain of the 
+parameters' variation;
+	\item knowing \textit{a priori} the conditional global minimizer of the objective 
+function;
+	\item generating the unconditional global minimizer of the objective function out of the 
+feasible domain;
+	\item generating the problems with a conditional minimum point located at the boundary 
+of the feasible domain;
+	\item defining the number of constraints active at the optimum point.
 \end{itemize}
 
-The demonstration of the developed approach in application to well-known index method for solving complex multiextremal optimization problems with non-convex constraints is considered.
+The demonstration of the developed approach in application to well-known index method for 
+solving complex multiextremal optimization problems with non-convex constraints is 
+considered.
 
-The developed approach allows generating any number of test global optimization problems with non-convex constraints for performing multiple computational experiments in order to obtain a reliable evaluation of the efficiency of the developed optimization algorithms.
+The developed approach allows generating any number of test global optimization problems 
+with non-convex constraints for performing multiple computational experiments in order to 
+obtain a reliable evaluation of the efficiency of the developed optimization algorithms.
 

### 2018-04-27 2018/LeGo/LeGo.tex
Some corrections

@@
 \usepackage{rotating}
-\usepackage{subfig}
 \usepackage{graphicx}
-\usepackage{tabularx}
 \usepackage{enumitem}
@@
 problems with a conditional minimum located at the boundary of the feasible domain and of 
-specifying the number of constraints active in the minimum points have been added. The 
+specifying the number of constraints active in the minimizer have been added. The 
 demonstration of the developed approach is given on the example of a known index method for 
-solving the complex multiextremal optimization problems with nonlinear constraints.
+solving the complex multiextremal optimization problems with non-convex constraints.
 \end{abstract}
@@
 In the present work, further development of the approach to generating the global optimization 
-problems with nonlinear constraints proposed originally in Ref.~\cite{Gergel2017} has been 
+problems with non-convex constraints proposed originally in \cite{Gergel2017} has been 
 conducted.
@@
 of the feasible domain relative to the whole global search domain can be specified. 
-At that, the location of the global minimum is known for the generated problems that simplifies 
-evaluating the results of the numerical experiments essentially. 
-in addition to the specified options, 
-the following functions have been added in the current version of the generator:
+At that, the location of the global minimum is known for the generated problems that essentially simplifies 
+evaluating the results of the numerical experiments. 
+In addition to the specified options, the following functions have been added in the current version of the generator:
 
@@
 For example, one generator for random samples of two-dimensional test functions has been 
-described in \cite{Grishagin2015}. 
-%This generator produces two-dimensional functions according to the formula
-%\begin{equation} \label{VAG}
-%\varphi(y)= -
-%\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2 + 
-%\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},
-%\end{equation}
-%where $g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2)$, $h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2)$, 
-%$y=(y_1,y_2)\in R^2, \; 0 \leq y_1,y_2 \leq 1$, and coefficients $A_{ij}, B_{ij}, C_{ij}, 
-%D_{ij}$  are taken uniformly in the interval $[-1,1]$.
+described in \cite{Grishagin2015}. %more refs
 Another generator (\textit{GKLS generator}) for the functions of arbitrary dimensionality with 
@@
 minimizer, etc.) has been proposed in \cite{Gaviano2003}.  Application of this generator for 
-studying some unconstrained optimization algorithms has been described in 
-\cite{Paulavicius2014,Sergeyev2015}.
+studying some unconstrained optimization algorithms has been described in \cite{Paulavicius2014,Sergeyev2015}.
 
@@
 generate the test global optimization problems with $m$ constraints based on the functions of 
-this class has been proposed in Ref.~\cite{Gergel2017}. In the present paper, the rules, which 
+this class has been proposed in \cite{Gergel2017}. In the present paper, the rules, which 
 allow formulating the constrained global optimization problems have been proposed so that:
@@
 property featuring the applied constrained optimization problems. 
-When generating the problems according to the scheme described in Ref.~\cite{Gergel2017}, 
+When generating the problems according to the scheme described in \cite{Gergel2017}, 
 the conditional minimum, in general, will be located inside the feasible domain. Therefore, the 

### 2018-04-29 2018/LeGo/LeGo.tex
Some corrections

@@
 
-For example, one generator for random samples of two-dimensional test functions has been 
-described in \cite{Grishagin2015}. %more refs
+A generator for random samples of two-dimensional test functions has been 
+described, for example, in \cite{Grishagin1997,Grishagin2015}. %more refs
 Another generator (\textit{GKLS generator}) for the functions of arbitrary dimensionality with 
@@
 The generator GCGen (Global Constrained optimization problem Generator) which allows to 
-generate the test global optimization problems with $m$ constraints based on the functions of 
-this class has been proposed in \cite{Gergel2017}. In the present paper, the rules, which 
-allow formulating the constrained global optimization problems have been proposed so that:
+generate the test global optimization problems with $m$ constraints has been proposed in \cite{Gergel2017}.
+In this paper, the rules, which allow formulating the constrained global optimization problems have been proposed so that:
 \begin{itemize}
@@
 \item From the point $\overline{y}$, a coordinate search of a feasible point $y^*$, for which at 
-least one constraint is active i.e. there is such index $j, 1 \leq j\leq m$ that $\left|g_j(y^*)\right| 
-\leq \delta$ is performed with given step $h$. 
+least one constraint is active (i.e. there is such index $j, 1 \leq j\leq m$ that $\left|g_j(y^*)\right| 
+\leq \delta$) is performed with given step $h$. 
 \item A rough estimate of $y^*$ obtained is refined by a local method. 
@@
 In the case, when it is necessary to generate a problem with a predefined number of active 
-constraints $S$ in the constrained minimum point, the rules are modified as follows.
+constraints $S$ in the minimizer, the rules are modified as follows.
 \begin{enumerate}
@@
 
-As an illustration, the level lines of the objective functions and the zero-level lines of four 
+As an illustration, the level lines of the objective functions and the zero-level lines of three 
 constraints for the problem constructed on the base of functions from \cite{Grishagin2015} with the volume 
@@
 \item A modification of the problem from item (b): constrained optimizer is located at the 
-boundary of the feasible domain, the number of active constraints is not defined.
+boundary of the feasible domain, the number of active constraints is not predefined.
 \item A modification of the problem from item (c): the number of active constraints is defined 
@@
 parameters' variation;
-	\item knowing \textit{a priori} the conditional global minimizer of the objective 
+	\item knowing a priori the conditional global minimizer of the objective 
 function;
@@
 feasible domain;
-	\item generating the problems with a conditional minimum point located at the boundary 
+	\item generating the problems with a conditional minimizer located at the boundary 
 of the feasible domain;
@@
 
-The demonstration of the developed approach in application to well-known index method for 
-solving complex multiextremal optimization problems with non-convex constraints is 
-considered.
-
 The developed approach allows generating any number of test global optimization problems 
-with non-convex constraints for performing multiple computational experiments in order to 
+with non-convex constraints for performing computational experiments in order to 
 obtain a reliable evaluation of the efficiency of the developed optimization algorithms.

### 2018-04-29 2018/Journal of Supercomputing/JoS_Paper.tex
Short version

@@
 
-\title{Parallel Algorithm for Solving Constrained Global Optimization Problems
+\title{Parallel Computing for Constrained Global Optimization Problems
 \thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}
@@
 \institute{K. Barkalov  \at
-               Institute of Information Technology, Mathematics and Mechanics\\ Lobachevsky State University of Nizhni Novgorod\\ Nizhni Novgorod, Russia \\
+               Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
               \email{konstantin.barkalov@itmm.unn.ru} 
@@
            I. Lebedev \at
-					Institute of Information Technology, Mathematics and Mechanics\\ Lobachevsky State University of Nizhni Novgorod\\ Nizhni Novgorod, Russia \\
+					Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
 							\email{ilya.lebedev@itmm.unn.ru} 
@@
 The \textit{index} $\nu(x)$ can also be defined by the conditions
-\[
+\begin{equation}\label{nu}
 g_i(y(x)) \leq 0, \; 1 \leq i < \nu, \; g_\nu(y(x))>0,
-\]
+\end{equation}
 where the last inequality is inessential if $\nu=m+1$.
@@
 	\item Determine the \textit{image} $y^k=y(x^k)$ in accordance with the mapping $y(x)$;
-	\item Compute the values $g_1(y^k),...,g_\nu(y^k),$ where the index $\nu \leq m$ is determined by the conditions
-	\[
-		g_i(y^k )\leq 0, \; 1 \leq i < \nu, \; g_\nu(y^k)>0, \; \nu \leq m.
-	\]
+	\item Compute the values $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from (\ref{nu}). 
 \end{itemize}
-The occurrence of the first violation of the constraint terminates the trial.
-In the case, when the point $y^k$ is a feasible one, i.e.
-when $x^k \in Q_{m+1},$ the trial includes the computation of the values of all problem functions and the index is accepted to be $\nu=m+1$.
-The pair of values
-\[
-\nu=\nu(x^k), \; z^k=g_\nu(y(x^k))
-\]
-is a \textit{trial result}.
+The pair of values $ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} $ is a \textit{trial result}.
 
@@
 
-The proposed parallelization scheme was implemented with the use of MPI technology.
-Main features of implementation consist in the following.
-A separate MPI-process is created for each of $S$ one-dimensional problems being solved, usually, one process per one processor employed.
-Each process can use $p$ threads, usually one thread per an accessible core.
+The proposed parallelization scheme was implemented with the use of MPI and OpenMP technologies.
+A separate MPI-process is created for each of $S$ one-dimensional problems being solved, one process per one processor employed.
+Each process use $p$ threads for solving one-dimensional problem, one thread per an accessible core.
 
-At every iteration of the method, the process with the index $s, 0\leq s < S,$ performs $p$ trials in parallel at the points $x^{s+iS}, 0\leq i < p$.
+%At every iteration of the method, the process with the index $s, 0\leq s < S,$ performs $p$ trials in parallel at the points $x^{s+iS}, 0\leq i < p$.
 At that, each process stores all $S\cdot p$ points, and an attribute indicating whether this point is blocked by another process or not is stored for each point.
@@
 
-At every iteration of the algorithm, operating within the $s$-th process, determines the coordinates of $p$ ``its own'' trial points.
+%At every iteration of the algorithm, operating within the $s$-th process, determines the coordinates of $p$ ``its own'' trial points.
 Then, the interchange of the coordinates of images of the trial points $y^{s+iS},\; 0\leq i<p,\; 0\leq s< S$ is performed (from each process to each one).
@@
 
-GKLS generator for the functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.) has been proposed in \cite{Gaviano2003}.
+%GKLS generator for the functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.) has been proposed in \cite{Gaviano2003}.
 Four GKLS classes of differentiable test functions of the dimensions $N = 4$ and $5$, have been used.
@@
 
-In this study we will use GKLS generator to produce the constrained problems.
+In this study we will use GKLS generator of the functions of arbitrary dimensionality proposed in \cite{Gaviano2003} to produce the constrained problems.
 The scheme that allows to form the constrained global optimization problems is proposed in \cite{Gergel2017}.
@@
 The problem was considered to be solved, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^*\right\|\leq \delta$.
-The size of the vicinity was selected as $\delta=0.03\left\|b-a\right\|$, where $a$ and $b$ are borders of the search domain.
+The size of the vicinity was selected as $\delta=0.03\left\|b-a\right\|$, where $a$ and $b$ are boundaries of the search domain.
 For the purpose of simulation of the computational complexity inherent to applied problems of optimization, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (series summation of $80$ thousand elements).
@@
 The data transfer overheads will not affect the total computational costs predominately in this case, and the time speedup will not differ from the iteration one so strongly.
+%Описание инфраструктуры
 
@@
 
-\bibitem{Paulavicius2014}
-Paulavi\v cius, R., Sergeyev, Y.D., Kvasov, D.E., \v Zilinskas, J. Globally-biased \textsc{Disimpl} algorithm for expensive global optimization. J. Glob. Optim. 59(2), 545--567 (2014)
+%\bibitem{Paulavicius2014} Paulavi\v cius, R., Sergeyev, Y.D., Kvasov, D.E., \v Zilinskas, J.
+Globally-biased \textsc{Disimpl} algorithm for expensive global optimization.
+J.
+Glob.
+Optim.
+59(2), 545--567 (2014)
 
-\bibitem{Sergeyev2015}
-Sergeyev, Y.D., Kvasov, D.E.: A deterministic global optimization using smooth diagonal auxiliary functions. Communications in Nonlinear Science and Numerical Simulation. 21(1-3), 99–111 (2015)
+%\bibitem{Sergeyev2015} Sergeyev, Y.D., Kvasov, D.E.: A deterministic global optimization using smooth diagonal auxiliary functions.
+Communications in Nonlinear Science and Numerical Simulation.
+21(1-3), 99–111 (2015)
 

### 2018-04-30 2018/LeGo/LeGo.tex
Исправил non- convex на non-convex, дважды :)

@@
 
-In the present paper, the methods for generating the global optimization test problems with non-
-convex constraints
+In the present paper, the methods for generating the global optimization test problems with non-convex constraints
 \begin{eqnarray}
@@
 
-This paper considers the method for generating the global optimization test problems with non-
-convex constraints, which allows:
+This paper considers the method for generating the global optimization test problems with non-convex constraints, which allows:
 \begin{itemize}

### 2018-05-03 2018/Journal of Supercomputing/JoS_Paper.tex
before translation

@@
 % Insert the name of "your journal" with
-% \journalname{myjournal}
+ \journalname{The Journal of Supercomputing}
 %
@@
 
+\bibitem{RefFerreiro}
+Ferreiro, A.M., Garcia, J.A., Lopez-Salas, J.G., Vazquez, C.: An efficient implementation of parallel simulated annealing algorithm in GPUs. J. Glob. Optim. 57(3), 863--890 (2013)
+
+\bibitem{RefZhu}
+Zhu, W.: Massively parallel differential evolution--pattern search optimization with graphics hardware acceleration: an investigation on bound constrained optimization problems.
+J.
+Glob.
+Optim.
+50(3), 417--437 (2011).
+
+\bibitem{Korosec}
+Koro\v sec, P., Vajter\v sic, M.,  \v Silc, J., Kutil, R.: Multi-core implementation of the differential ant-stigmergy algorithm for numerical optimization. J. Supercomput. 63(3), 757--772 (2013)
+
+\bibitem{Guerrero}
+Guerrero, G.D., Cecilia, J.M., Llanes, A. et al.: Comparative evaluation of platforms for parallel Ant Colony Optimization. J. Supercomput. 69(1) 318--329 (2014)
+
 \bibitem{Jones2001}

### 2018-05-04 2018/Journal of Supercomputing/JoS_Paper.tex
Before proofreading

@@
 \begin{abstract}
-This work considers a parallel algorithm for solving multiextremal problems with non-convex constraints.
-The distinctive feature of this algorithm, which does not use penalty functions, is the separate consideration of each problem constraint.
-The search process can be conducted by reducing the original multidimensional problem to a number of related one-dimensional problems and solving this set of problems in parallel.
-An experimental assessment of parallel algorithm efficiency was conducted by finding the numeric solution to several hundred randomly generated multidimensional multiextremal problems with non-convex constraints.
+In the present work, a parallel algorithm for solving the multiextremal problems with non-convex constrains is considered.
+Within the framework of approach considered, the solving of initial multidimensional problem is reduced to the parallel solving of a set of linked one-dimensional problems.
+The parallelization scheme is featured by the absence of a single control process since the solution of any one-dimensional problem is a solution of the initial multidimensional one.
+The non-convex constraints are accounted for by a novel index scheme not utilizing the ideas of penalty functions.
+The efficiency of the parallel algorithm was confirmed experimentally by the numerical solving of several hundred multidimensional multiextremal problems with non-convex constrains generated randomly.
 \keywords{Global optimization \and Constrained problems \and Non-convex constraints \and Dimensionality reduction \and Parallel algorithms}
@@
 Moreover, it is supposed that even a single computation of the function value can be time-consuming, as in the applied problems it requires performing numerical simulation.
-These assumptions are typical for many approaches to building parallel algorithms for unconstrained global optimization \cite{Jones2001,Paulavicius2011,Evtushenko2013}.
+%These assumptions are typical for many approaches to building parallel algorithms for unconstrained global optimization \cite{Jones2001,Paulavicius2011,Evtushenko2013}.
 
-At the same time, it is common for applied constrained optimization problems to be in a situation where violating one constraint results in all other functions returning indeterminate values.
-An example includes optimal control problems, described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2011,Balandin2017}.
+The numerical algorithms for solving the problems of the specified class are featured by an exponential growth of the computational costs with increasing dimensionality (so called ``curse of dimensionality'').
+The use of supercomputers expands the field of application of the global optimization methods and, at the same time, sets a problem of efficient parallelization of the search process.
+
+At present, the parallel algorithms, which utilize the heuristic approaches and based on the idea of random search in any way are quite popular (see \cite{RefFerreiro,RefZhu,Korosec,Guerrero}).
+However, because of the probabilistic nature, the algorithms of such type guarantee the convergence to the global optimum solution in the probabilistic meaning only, that distinguishes these ones unfavorably from the deterministic parallel global optimization algorithms \cite{Jones2001,Paulavicius2011,Evtushenko2013}.
+The efficient deterministic algorithms are oriented onto the solving of the unconstrained optimization problems as a rule.
+The problems with constraints are solved using the penalty function method usually having a number of disadvantages.
+One of these is the impossibility of its application in the problems with the \textit{partially defined functions} (i.e.
+when violating any constraint, the values of the rest functions of the problem remain undefined).
+
+As the examples, one can mention the optimal control problems, described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2011,Balandin2017}.
 It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e.
@@
 
-This partial computability of the functions in constrained optimization problems substantially complicates the application of the well-known penalty function method (in some cases making it completely impossible).
-Thanks to its simplicity, this method is one of the most popular approaches to solving problems with constraints.
-However, calculating the penalty function requires first calculating the values of all of the problem’s functions at the given point, which is impossible if they are partially indeterminate.
-
-In this work, the authors consider an approach to minimizing multiextremal functions under non-convex constraints, developed in \cite{Strongin2000,Sergeyev2001,Barkalov2002} and called \textit{the index method}.
-The approach is based on separate considering every constraint in the problem and is not related to using penalty functions.
-According to the index method, each iteration (\textit{a trial}) at a respective point in the search domain includes a sequential check of the problem constraints at that point.
-As soon as the first constraint violation is found, the trial is interrupted and the method proceeds with the next iteration; no other problem functions are calculated at that point.
-This allows problems to be solved in which function values may not be determined for the entire search domain.
-Under this approach, solving multidimensional problems is reduced (using Peano-type space-filling curves) to solving equivalent one-dimensional problems.
-
-It should be noted that standard approaches to algorithm parallelization are not quite applicable to global optimization.
-For example, the rules for selecting another iteration point are quite simple and do not require parallelization (as overheads associated with organizing parallel computations will nullify any possible acceleration).
-Some acceleration can be achieved by parallelizing the computation of function values describing the object to be optimized; however, this approach is specific to each individual problem being solved.
-
-The following approach looks more promising.
-The algorithm can be modified to run several trials in parallel.
-This approach provides the efficiency (as parallelization is applied to the most computation-intensive part of the problem solving process) and generality (in that it applies to a wide range of global optimization algorithms).
-The approach, described in \cite{Sidorov2015} for unconstrained optimization, was used in this work for parallelizing constrained optimization algorithms.
+In the present work, a novel approach to the minimization of the multiextremal functions with non-convex constraints developed in \cite{Strongin2000,Sergeyev2001,Barkalov2002} and referred to as the index method  of accounting for the constraints was applied.
+The approach is based on a separate accounting for each constraint of the problem and is not related to the use of the penalty functions.
+The solving of initial multidimensional problem is reduced (using the Peano-type space-filling curves) to solving an equivalent set of the one-dimensional problems, which can be solved in parallel in such a way that each of these processes shares the information obtained by the other processes.
+The propose scheme was implemented using Lobachevsky supercomputer with the use of the distributed memory as well as the shared one.
 
@@
 with a priory unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal.
-It is admitted that functions $g_i(y)$ are defined and computable only at the points $y \in D$ satisfying the conditions
+It is assumed that functions $g_i(y)$ are defined and computable only at the points $y \in D$ satisfying the conditions
 \begin{equation}\label{g_k}
@@
 For the purpose of simulation of the computational complexity inherent to applied problems of optimization, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (series summation of $80$ thousand elements).
+
+Computational experiments were carried out on a high-performance cluster of Lobachevsky State University of Nizhni Novgorod.
+The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.
+The CPU had 8 cores, i.e.
+16 cores were available in each node.
 

### 2018-05-07 2018/PAVT_Toropov/PAVT2018_MAM.tex
After checking

--- 2018-01-24 2018/PAVT_Toropov/PAVT2018_MAM.tex
@@
 % first the title is needed
-\title{Parallel Multipoint Approximation Method for Large Scale Optimization Problems 
-\thanks{This study was supported by the Russian Science Foundation, project No. 16-11-10150.}}
+\title{Parallel Multipoint Approximation Method for Large-Scale Optimization Problems\thanks{This
+study was supported by the Russian Science Foundation, project
+No.\,16-11-10150.}}
 
@@
 
-\author{ Victor P. Gergel%(ORCID 0000-0002-4013-2329)
+\author{Victor P. Gergel%(ORCID 0000-0002-4013-2329)
 $^1$ \and
@@
 \begin{abstract}
-The paper presents a new development in the Multipoint Approximation Method (MAM) that makes it capable of handling large scale problems.
-The approach relies on approximations built in the space of design variables within the iterative trust-region-based framework of MAM.
-With the purpose of solving the problems of high dimensionality in a reasonable time, a parallel variant of the Multipoint Approximation Method (PMAM) has been developed.
-It was supposed that the values of the objective function and of the constraints were computed using the distributed memory (on several cluster nodes), whereas the optimization module run on a single node using the shared memory.
-The numerical experiments have been carried out on a benchmark example of structural optimization.
+The paper presents a new development in the Multipoint Approximation Method
+(MAM) that makes it capable of handling large-scale problems. The approach
+relies on approximations built in the space of design variables within the
+iterative trust-region-based framework of MAM. With the purpose of solving high
+dimensionality problems in a reasonable time, a parallel variant of the
+Multipoint Approximation Method (PMAM) has been developed. It is supposed that
+the values of the objective function and those of the constraints are computed
+using distributed memory (on several cluster nodes), whereas the optimization
+module runs on a single node using shared memory. Numerical experiments have
+been carried out on a benchmark example of structural optimization.
 
-\keywords{design optimization, multidisciplinary optimization, multipoint approximation method, parallel computing.}
+\keywords{design optimization, multidisciplinary optimization, multipoint
+approximation method, parallel computing.}
 \end{abstract}
@@
 
-In the present paper, the multipoint approximation method (MAM \cite{Toropov1989,Toropov1992,ToropovFilatov1993}) and its application to the large scale optimization problems are considered.
-In problems with a large (in the order of hundreds) number of design variables MAM proved to be efficient, e.g.
-in turbomachinery applications \cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010}.
-This method is an iterative optimization technique based on mid-range approximations built in trust regions.
-A trust region is a sub-domain of the design space in which a set of design points, produced according to a small-scale design of experiments (DoE), are evaluated.
-These and a subset of previously evaluated design points are used to build metamodels of the objective and constraint functions that are considered to be valid within a current trust region.
-The trust region will then translate and change size as optimization progresses.
-The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values \cite{KeulenToropovMarkine1996,ToropovKeulenMarkine1996} and occasional simulation failures \cite{ToropovMarkineHolden1999}.
-The mid-range approximations used in the trust regions, as originally suggested in \cite{Toropov1989} for structural optimization problems, are intrinsically linear functions (i.e.
-nonlinear functions that can be led to a linear form by a simple transformation) for individual sub-structures, and assembly of them for the whole structure.
-This was enhanced by the use of gradient-assisted metamodels \cite{ToropovFilatov1993}, use of simplified numerical models that is also termed a multi-fidelity approach,\cite{ToropovMarkine1996} and the use of analytical models derived by Genetic Programming \cite{ToropovAlvarez1998}.
-One of the recent developments \cite{PolynkinToropov2012} involved the use of approximation assemblies, i.e.
-a two stage approximation building process that is conceptually similar to the original one used in \cite{Toropov1989} but is free from the limitation that lower level approximations are linked to individual substructures.
+In the present paper, the multipoint approximation method (MAM)
+\cite{Toropov1989,Toropov1992,ToropovFilatov1993} and its application to
+large-scale optimization problems are considered. In problems with a large (in
+the order of hundreds) number of design variables, MAM has proved to be
+efficient, e.g., in turbomachinery applications
+\cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010}.
+This method is an iterative optimization technique based on mid-range
+approximations built in trust regions. A trust region is a subdomain of the
+design space in which a set of design points, produced according to a
+small-scale design of experiments (DoE), is evaluated. These and a subset of
+previously evaluated design points are used to build metamodels of the
+objective and constraint functions that are considered to be valid within a
+current trust region. The trust region will then translate and change size as
+optimization progresses. The trust region strategy has gone through several
+stages of development to account for the presence of numerical noise in the
+response function values
+\cite{KeulenToropovMarkine1996,ToropovKeulenMarkine1996} and occasional
+simulation failures \cite{ToropovMarkineHolden1999}. The mid-range
+approximations used in the trust regions, as originally suggested in
+\cite{Toropov1989} for structural optimization problems, are intrinsically
+linear functions (i.e. nonlinear functions that can be reduced to a linear form
+by a simple transformation) for individual substructures, and an assembly of them
+for the whole structure. This was enhanced by the use of gradient-assisted
+metamodels \cite{ToropovFilatov1993}, the use of simplified numerical models
+which is also termed the multi-fidelity approach \cite{ToropovMarkine1996}, and
+the use of analytical models derived by genetic programming
+\cite{ToropovAlvarez1998}. One of the recent developments
+\cite{PolynkinToropov2012} involves the use of approximation assemblies, i.e. a
+two stage approximation building process that is conceptually similar to the
+original one used in \cite{Toropov1989} but is free from the limitation that
+lower level approximations are linked to individual substructures.
 
-The Moving Least Squares Method (MLSM) was proposed in \cite{LancasterSalkauskas1981} for smoothing and interpolation of scattered data and later used in the mesh-free form of the finite element method (FEM) \cite{Liszka1984}.
-As suggested in \cite{ChoiYounYang2001}, it can be used as a technique for metamodelling and used in multidisciplinary optimization (MDO) frameworks.
-The MLSM is a weighted least squares method where the weights depend on the Euclidean distance from a sample point to where the surrogate model is to be evaluated.
-The weight value for a certain sample point decays as the distance increases.
-Describing the weight decay with a Gaussian function tends to be the most useful option even though many others have been evaluated in \cite{ToropovSchrammSahaiJones2005}.
-As demonstrated in \cite{PolynkinToropov2010}, the cross-validated MLSM can be used both for design variable screening and for surrogate modelling.
-In order to create an efficient MDO framework for problems with disparate discipline attributes \cite{OllarToropovJones2014} extended the optimization approach of MAM to the use of local DOEs and MLS approximations built in different subspaces of the total design variable space corresponding to the individual disciplines.
-The subspaces are finally combined into the total design variable space in which the resulting MDO problem is solved.
+The Moving Least-Squares Method (MLSM) was proposed in
+\cite{LancasterSalkauskas1981} for smoothing and interpolation of scattered
+data and was later used in the mesh-free form of the finite element method
+(FEM) \cite{Liszka1984}. As suggested in \cite{ChoiYounYang2001}, it can be
+used as a technique for metamodeling and in multidisciplinary optimization
+(MDO) frameworks. The MLSM is a weighted least-squares method where the weights
+depend on the Euclidean distance from a sample point to where the surrogate
+model is to be evaluated. The weight value for a certain sample point decays as
+the distance increases. Describing the weight decay with a Gaussian function
+tends to be the most useful option, even though many others have been evaluated
+in \cite{ToropovSchrammSahaiJones2005}. As demonstrated in
+\cite{PolynkinToropov2010}, the cross-validated MLSM can be used both for
+design variable screening and for surrogate modeling. In order to create an
+efficient MDO framework for problems with disparate discipline attributes, the
+optimization approach of MAM was extended  in \cite{OllarToropovJones2014} to
+the use of local DOEs and MLS approximations built in different subspaces of
+the total design variable space corresponding to the individual disciplines.
+The subspaces are finally combined into the total design variable space in
+which the resulting MDO problem is solved.
 
-This paper presents a Parallel Multipoint Approximation Method that makes it capable of handling problems with the number of design variables in the order of thousands.
-The parallel variant of the algorithm (with the use of the shared memory) has been developed with the purpose of minimizing the worktime of the part of the algorithm related to constructing the approximation and solving the approximated problem and not related to computing the values of the objective function and  constraints.
-The processes of computing the values of the objective function and constraints (with the use of the
-distributed memory) are supposed to be parallelized already.
+This paper presents a Parallel Multipoint Approximation Method that makes it
+capable of handling problems with numbers of design variables in the order of
+thousands. The parallel variant of the algorithm (with the use of shared
+memory) has been developed with the purpose of minimizing the work time of the
+part of the algorithm related to constructing the approximation and solving the
+approximated problem, but not related to computing the values of the objective
+function and constraints. The processes of computing the values of the
+objective function and constraints (with the use of distributed memory) are
+supposed to be already parallelized.
 
@@
 
-It is useful to start with a brief description of MAM. A typical formulation of a constrained optimization problem that MAM works with is as follows:
+It would be useful to start with a brief description of MAM. A typical
+formulation of a constrained optimization problem that MAM works with is as
+follows:
 \begin{equation}
@@
   \min\limits_{a_i \le x_i \le b_i}F_0(x) \\
-  s.t.\; F_j(x) \le 1,\; j=1,\dots ,M,
+  \text{s.t.}\ F_j(x) \le 1,\ j=1,\dots ,M,
   \end{array}
 \end{equation}
-where $x$ is the vector of design variables, $a$ and $b$ are the lower and upper bounds for the design variables, respectively, $F_0(x)$ is the objective function, and $F_j(x)$ are the constraints.
-The numbers of design variables and constraints are $n$ and $M$, respectively.
-MAM attempts to solve this problem by using approximations of the objective function and constraints in a series of trust regions.
-The trust region strategy seeks to zoom in on the region where the constrained minimum is achieved.
-It aims at finding a trust region that is sufficiently small for the approximations to be of sufficiently good quality to improve the design, and that contains the point of the constrained minimum as its interior point.
-The main loop of the MAM is organized as follows.
-\\
+where $x$ is a vector of design variables, $a$ and $b$ are the lower and upper
+bounds for the design variables, respectively, $F_0(x)$ is the objective
+function, and $F_j(x)$ are the constraints. The numbers of design variables and
+constraints are $n$ and $M$, respectively. MAM attempts to solve this problem
+by using approximations of the objective function and constraints in a series
+of trust regions. The trust region strategy seeks to zoom in on the region
+where the constrained minimum is achieved. It aims at finding a trust region
+that is sufficiently small for the approximations to be of sufficiently good
+quality to improve the design and contains the point of the constrained minimum
+as an interior point. The main loop of the MAM is organized as follows.
 
-\textbf{Algorithm (MAM).}
+\subsection*{Algorithm (MAM).}
 \begin{enumerate}
-\item Initialization: choose a starting point $ x^0$ and initial trust region $[ a^0,  b^0]$ such that $ x^0 \in [ a^0,  b^0]$.
-\item At the $k$-th iteration the current approximation to the constrained minimum is $ x^k$, the current trust region is $[ a^k,  b^k] \subset [ a^0,  b^0]$.
+\item Initialization: choose a starting point $ x^0$ and initial trust
+    region $[ a^0,  b^0]$ such that $ x^0 \in [ a^0,  b^0]$.
+
+\item On the $k$th iteration, the current approximation to the constrained
+    minimum is $ x^k$, and the current trust region is $[ a^k,  b^k]
+    \subset [ a^0,  b^0]$.
   \begin{enumerate}[label=(\alph*)]
-    \item Design of Experiments (DoE): a set of points $ x_k^i \in [ a^k,  b^k]$ is chosen to be used for approximation building.
-Responses are evaluated at the DoE points and approximations are built using the obtained values.
-Currently, the pool of approximation methods available in MAM consists of a metamodel assemblies \cite{PolynkinToropov2012} and the moving least-squares metamodels \cite{LancasterSalkauskas1981,Liszka1984,ChoiYounYang2001,ToropovSchrammSahaiJones2005}.
+    \item Design of Experiments (DoE). A set of points $ x_k^i \in [
+        a^k,  b^k]$ is chosen to be used for building approximations.
+        Responses are evaluated at the DoE points and approximations
+        are built using the obtained values. Currently, the pool of
+        approximation methods available in MAM consists of metamodel
+        assemblies \cite{PolynkinToropov2012} and the moving
+        least-squares metamodels
+        \cite{LancasterSalkauskas1981,Liszka1984,ChoiYounYang2001,ToropovSchrammSahaiJones2005}.
         Other metamodel types could be used as well.
@@
     Denote the approximate objective function and constraints by $\widetilde{F}^k_0(x)$ and $\widetilde{F}^k_j(x)$, respectively.
-    \item The original optimization problem (\ref{eq:problem}) is replaced by the following problem:
+
+    \item The original optimization problem (\ref{eq:problem}) is
+        replaced by the following:
     \begin{equation}
@@
       \min\limits_{a_i^k \le x_i ^k\le b_i}\widetilde{F}^k_0(x) \\
-      s.t.\; \widetilde{F}^k_j( x) \le 1,\; j=1,\dots ,M.
+      \text{s.t.}\ \widetilde{F}^k_j( x) \le 1,\ j=1,\dots ,M.
       \end{array}
     \end{equation}
-    The approximate problem (\ref{eq:problem_approx}) is solved using Sequential Quadratic Programming (SQP) and the solution of this problem determines the center of the next trust region.
-    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $ x^k$, and on the size of the current trust region \cite{KeulenToropovMarkine1996}.
-    \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $ x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations).
-If the termination criterion is satisfied, the algorithm proceeds to step 3.
-Otherwise, it returns to step 2.
+    The approximate problem (\ref{eq:problem_approx}) is solved using
+    Sequential Quadratic Programming (SQP). The solution of this
+    problem determines the center of the next trust region.
+
+    \item The size of the next trust region is determined depending on
+        the quality of approximations on the previous iteration, on the
+        history of points $ x^k$, and on the size of the current trust
+        region \cite{KeulenToropovMarkine1996}.
+
+    \item The termination criterion is checked (it is a part of the
+        trust region strategy and depends on the position of the point
+        $ x^{k+1}$ in the current trust region, the size of the current
+        trust region and the quality of approximations). If the
+        termination criterion is satisfied, the algorithm proceeds to
+        step 3. Otherwise, it returns to step 2.
   \end{enumerate}
-  \item Optimization terminates. The obtained approximation to the solution of the problem (\ref{eq:problem}) is $ x^{k+1}$.
+  \item Optimization terminates. The obtained approximation to the solution
+      of problem (\ref{eq:problem}) is $ x^{k+1}$.
 \end{enumerate}
 
-The selection of the approximations $\widetilde{F}^k_j(x),\; j=0, \dots ,M,$ is such that their evaluation is inexpensive as compared to the evaluation of the original response functions $F_j(x)$.
-For example, intrinsically linear functions were successfully used for a variety of design optimization problems in the works \cite{ToropovFilatov1993,KeulenToropov1997}.
-The approximations are determined by means of the weighted least squares:
+The approximations $\widetilde{F}^k_j(x),\; j=0, \dots ,M,$ are selected in
+such a way that their evaluation is inexpensive as compared to the evaluation
+of the original response functions $F_j(x)$. For example, intrinsically linear
+functions were successfully used for a variety of design optimization problems
+in \cite{ToropovFilatov1993,KeulenToropov1997}. The approximations are
+determined by means of the weighted least squares:
 \begin{equation}\label{eq:least_sq}
@@
 \end{equation}
-In (\ref{eq:least_sq}) minimization is carried out with respect to the tuning parameters $a_j$; $w_{pj}$ are the weight coefficients, and $P$ is the number of sampling points in Design of Experiments (DoE) which must not be less than the number of parameters in the vector $a_j$.
+In (\ref{eq:least_sq}), minimization is carried out with respect to the tuning
+parameters $a_j$; $w_{pj}$ are the weight coefficients, and $P$ is the number
+of sampling points in Design of Experiments (DoE), which must not be less than
+the number of parameters in the vector $a_j$.
 
-The weight coefficients $w_{pj}$ strongly influence the difference in the quality of the approximations in different regions of the design variable space.
-Since in realistic constrained optimization problems the optimum point usually belongs to the boundary of the feasible region, the approximation functions should be more accurate in such domain.
-Thus, the information at the points located near the boundary of the feasible region is to be treated with greater weights.
-In a similar manner a larger weight can be allocated to a design with a better objective function, see \cite{ToropovFilatov1993,KeulenToropov1997}.
+The weight coefficients $w_{pj}$ strongly influence the difference in the
+quality of the approximations in different regions of the design variable
+space. Since in realistic constrained optimization problems the optimum point
+usually belongs to the boundary of the feasible region, the approximation
+functions should be more accurate in such domain. Thus, the information at the
+points located near the boundary of the feasible region is to be treated with
+greater weights. In a similar manner, a larger weight can be allocated to a
+design with a better objective function (see
+\cite{ToropovFilatov1993,KeulenToropov1997}).
 
-As optimization steps are carried out, a database with response function values becomes available.
-In order to achieve good quality approximations in the current trust region, an appropriate selection of DoE points must be made.
-In this work, DoE points in each trust region is generated randomly.
-Generally, points located far from the current trust region would not contribute to the improvement of the quality of the resulting approximations in the trust region.
-For this reason only points located in the neighborhood of the current trust region are taken into account, as depicted in Fig.~\ref{fig:trustregion}.
+As optimization steps are carried out, a database with response function values
+becomes available. In order to achieve good quality approximations in the
+current trust region, an appropriate selection of DoE points must be made. In
+this work, DoE points in each trust region are generated randomly. Generally,
+points located far from the current trust region would not contribute to the
+improvement of the quality of the resulting approximations in the trust region.
+For this reason, only points located in a neighborhood of the current trust
+region are taken into account, as depicted in Fig.\,\ref{fig:trustregion}.
 \begin{figure}[ht]
@@
     \includegraphics[width=0.5\textwidth]{trustregion.png}
-    \caption{Current trust region (smaller box) and its extension (larger box): points outside the larger box are not used for building the approximate functions}
+    \caption{Current trust region (smaller box) and its extension (larger box):
+    points outside the larger box are not used for building the approximate
+    functions}
     \label{fig:trustregion}
 \end{figure}
-A box in the space of design variables, which is approximately 1.5 to 1.8 times larger than the box representing the current trust region, was found by numerical experimentation to be a reasonable choice for the size of the neighborhood.
+A box in the space of design variables, which is approximately 1.5 to 1.8 times
+larger than the box representing the current trust region, was found by
+numerical experimentation to be a reasonable choice for the size of the
+neighborhood.
 
-In this work an approach is used that is based on the assembly of different approximate models $\{\varphi_l\}$ into one metamodel using the following form (note that the indices $j$ and $k$ are suppressed to simplify notation):
+In this work, an approach is used that is based on the assembly of different
+approximate models $\{\varphi_l\}$ into one metamodel using the following form
+(note that the indices $j$ and $k$ are suppressed to simplify notation):
 \begin{equation}\label{eq:assembly}
@@
 \end{equation}
-where $NF$ is the number of regressors in the model pool $\{\varphi_l\}$ and $b_l$ are corresponding regression coefficients.
-The used procedure consists of two subsequent steps.
-In the first step, the parameters $a_l$ of individual functions (regressors) $\varphi_l$ in (\ref{eq:assembly}) are determined by solving a weighted least squares problem using a specified a DoE of $P$ points:
+where $NF$ is the number of regressors in the model pool $\{\varphi_l\}$, and
+$b_l$ are the corresponding regression coefficients. The  procedure used
+consists of two subsequent steps. In the first step, the parameters $a_l$ of
+individual functions (regressors) $\varphi_l$ in (\ref{eq:assembly}) are
+determined by solving a weighted least-squares problem using a specified DoE of
+$P$ points:
 %\begin{equation}\label{eq:least_sq}
@@
 
-In the second step, based on the same DoE and keeping the obtained parameters $a_l$ fixed, a vector $b$ in (\ref{eq:assembly}) is estimated using the following formulation
+In the second step, based on the same DoE and keeping the obtained parameters
+$a_l$ fixed, a vector $b$ in (\ref{eq:assembly}) is estimated using the
+following formulation:
 \[
@@
 \]
-that leads to solving a linear system of $NF$ equations with $NF$ unknowns $b_l$, where $NF$ is the number of regressors in the model pool $\{\varphi_l\}$.
+which leads to solving a linear system of $NF$ equations with $NF$ unknowns
+$b_l$, where $NF$ is the number of regressors in the model pool
+$\{\varphi_l\}$.
 
-The selection of the regressors $\{\varphi_l\}$ is based on the number of sampling points currently located in the trust region.
-In the mid-range approximation framework, inexpensive approximate models for objective and constraint functions are built using minimum required number of sampling points.
-The simplest case is a linear function of the tuning parameters $a$:
+The selection of the regressors $\{\varphi_l\}$ is based on the number of
+sampling points currently located in the trust region. In the mid-range
+approximation framework, inexpensive approximate models for objective and
+constraint functions are built using the minimum required number of sampling
+points. The simplest case is that of a linear function of the tuning parameters
+$a$:
 \[
@@
 This structure can be extended to an \textit{intrinsically linear} function.
-Such functions are nonlinear, but they can be led to linear ones by simple transformations.
-The most useful function among them is the multiplicative function
+Such functions are nonlinear but they can be reduced to linear ones by simple
+transformations. The most useful function among them is the multiplicative
+function
 \[
@@
 
-Intrinsically linear functions have been successfully used for a variety of design optimization problems. 
-The advantage of these approximation functions is that a relatively small number $N+1$ ($N$ is number of design variables) of tuning parameters $a_i$ is to be determined, and the corresponding least squares problem is solved easily.
-This is the most important feature of such approximations as it allows applying them to large scale optimization problems.
+Intrinsically linear functions have been successfully used for a variety of
+design optimization problems. The advantage of these approximation functions is
+that a relatively small number $N+1$ ($N$ is the number of design variables) of
+tuning parameters $a_i$ is to be determined, and the corresponding
+least-squares problem is solved easily. This is the most important feature of
+such approximations as it allows applying them to large-scale optimization
+problems.
 
-Other intrinsically linear functions may be considered in the model pool, e.g.
+Other intrinsically linear functions may be considered in the model pool, e.g.,
 \[
@@
 
-As more points are added to the database the approximations may be switched to higher quality models, e.g. a rational model
+As more points are added to the database, the approximations may be switched to
+higher quality models, e.g., a rational model
 \begin{equation}\label{eq:rational}
-\varphi(x) = \frac{a_1 + a_2x_1 + a_3x_2+...+a_{n+1}x_n}{1 + a_{n+2}x_1 + a_{n+3}x_2+...+a_{2n+1}x_n},
+\varphi(x) = \frac{a_1 + a_2x_1 + a_3x_2+...+a_{n+1}x_n}{1 + a_{n+2}x_1 + a_{n+3}x_2+...+a_{2n+1}x_n}.
 \end{equation}
-The coefficients in (\ref{eq:rational}) are determined using a least squares approach which reduces to a nonlinear optimization problem with a constraint on the sign of the denominator (positive or negative).
-The latter is necessary in order to prevent the denominator from crossing the zero axis within a specified trust region.
-One may note that this formulation may yield the objective function with many local minima.
-Currently this problem is resolved using optimization restarts from a specified number of initial guesses randomly generated in a trust region.
+The coefficients in (\ref{eq:rational}) are determined using a least-squares
+approach which reduces to a nonlinear optimization problem with a constraint on
+the sign of the denominator (positive or negative). The latter is necessary in
+order to prevent the denominator from crossing the zero axis within a specified
+trust region. One may note that this formulation may yield an objective
+function with many local minima. Currently, this problem is resolved using
+optimization restarts from a specified number of initial guesses randomly
+generated in a trust region.
 
-Tests results demonstrated that, although the above functions may describe the global behavior rather poorly, such approximations proved to be efficient in the mid-range approximation framework of MAM.
+Tests results demonstrated that, although the above functions may describe the
+global behavior rather poorly, such approximations prove to be efficient in the
+mid-range approximation framework of MAM.
 
@@
 
-Let us consider possible methods of parallelization, which could be applied for the considered problems.
+Let us consider possible methods of parallelization that could be applied to
+the problems considered.
 
-First, one can parallelize the computing of the functions describing the optimized object.
-This way is obvious as well as necessary one since in the industrial design optimization problems the computing of even a single function value might take several hours.
-However, this method is a specific one for each particular problem.
-Here the issues of computations are addressed at the level of the application software, in which the industrial modeling is performed (e.g.
-Ansys, OpenFOAM, etc.).
+First, one can parallelize the computation of the functions describing the
+optimized object. This way is an obvious as well as necessary one since, in
+industrial design optimization problems, the computation of even a single
+function value may take several hours. However, this method is a specific one
+for each particular problem. Here the computation issues are addressed at the
+level of the application software in which the industrial modeling is performed
+(e.g., Ansys, OpenFOAM, etc.).
 
-Second, one can correct the algorithm with the purpose of parallel computing several values of the objective function and constraints at different points of the search domain.
-According to the MAM rules, in design of experiments $P$
-sampling points are formed in current trust region at every iteration.
-The function values at these points can be computed on different processors (nodes) in parallel.
-The above corresponds to the parallelization of Step 2a of the algorithm with the use of the distributed memory.
-The number of sampling points generated within each iteration may be set equal to $P = k\cdot NP$ number, where $NP$ is a number of available processors (or nodes) and $k \geq 1$.
-In  terms of the time, the latter will be equivalent to $NP$ function evaluations per step.
-This method has been implemented successfully in \cite{PolynkinToropov2008} and has demonstrated a good efficiency since at the number of the design variables of the order of 100 major time is consumed by the function evaluations, the work of the MAM itself (in the sequential regime) introduces minor overhead.
+Second, one can correct the algorithm with the purpose of parallelly computing
+several values of the objective function and constraints at different points of
+the search domain. According to the MAM rules, in design of experiments, $P$
+sampling points are formed in the current trust region at each iteration. The
+function values at these points can be computed on different processors (nodes)
+in parallel. The above corresponds to the parallelization of Step 2a of the
+algorithm using distributed memory. The number of sampling points generated
+within each iteration may be set equal to $P = k\cdot NP$, where $NP$ is the
+number of available processors (or nodes), and $k \geq 1$. In terms of time,
+the latter will be equivalent to $NP$ function evaluations per step. This
+method has been implemented successfully in \cite{PolynkinToropov2008} and has
+demonstrated a good efficiency since, for numbers of design variables of the
+order of 100, time is mainly consumed by the function evaluations, whereas the
+work of the MAM itself (in the sequential regime) introduces a minor overhead.
 
-However, when the number of variables becomes of the order of 1000, the work of the sequential part of MAM begins to affect the total problem solving time essentially (assuming the time of computing the objective function
-values remains constant).
-Thus, for the problem considered in Section 4 the time of execution of a single iteration of the method increased more than 4000 times (from 1.5 up to 6000 seconds) when increasing the number of variables from 100 up to  1000.
-Therefore, another approach to the parallelization of the algorithm has been applied within the framework of the
-present study: namely, the computational rules of MAM  providing the construction of the approximating, the solving of the approximated problem, and the choice of the next trust region (Steps 2b, 2c, and 2d of the algorithm) have been parallelized.
+However, when the number of variables becomes of the order of 1000, the work of
+the sequential part of MAM begins to affect the total problem solving time
+essentially (assuming that the time of computing the objective function values
+remains constant). Thus, for the problem considered in Section 4, the time of
+execution of a single iteration of the method increased by a factor of more
+than 4000 (from 1.5 up to 6000 seconds) when increasing the number of variables
+from 100 up to 1000. Another approach to the parallelization of the algorithm
+has therefore been applied within the framework of the present study: namely,
+the computational rules of MAM  providing for the construction of the
+approximation, the solution of the approximated problem, and the choice of the
+next trust region (Steps 2b, 2c, and 2d of the algorithm) were parallelized.
 
-In order to find the most time-consuming parts of the sequential program developed earlier, we have applied Intel VTune Amplifier XE.
-The performed analysis has shown the matrix multiplication and solving SLAE when constructing the approximations as well as solving the approximating problem by SQP method to be the most time-consuming operations in the execution of MAM.
-The matrix multiplication and solving SLAE are standard operations implemented in many high-performance libraries.
-Here we used the corresponding parallel methods from Intel MKL library.
-We have parallelized the SQP method ourselves using OpenMP.
+In order to find the most time-consuming parts of the sequential program
+developed earlier, we applied the Intel VTune Amplifier XE. The analysis
+performed has shown that the most time-consuming operations in the execution of
+MAM are matrix multiplication, the solution of the SLAE when constructing the
+approximations, and the solution of the approximating problem by SQP. Matrix
+multiplication and solution of SLAE are standard operations implemented in many
+high-performance libraries. Here we used the corresponding parallel methods
+from the Intel MKL library. We parallelized the SQP method ourselves using
+OpenMP.
 
@@
 
-The example considered in this study is a classical engineering optimization problem known as the scalable cantilevered beam \cite{Vanderpllaats2001}.
-The engineering object to be optimized is shown in Fig.
-\ref{fig:beam} taken from \cite{Vanderpllaats2001}.
+The example considered in this study is a classical engineering optimization
+problem known as the scalable cantilevered beam \cite{Vanderpllaats2001}. The
+engineering object to be optimized is shown in Fig.\,\ref{fig:beam} (taken from
+\cite{Vanderpllaats2001}).
 
@@
 The design variables are the widths $b_i$ and heights $h_i$ of the segments.
-The number of segments $N$ can be chosen arbitrary.
-The total length of the beam is 500 cm, the lengths of the segments are $l_i=500/N$  cm.
-There are $N$ geometric constraints (the aspect ratios of each block (i.e.
-heights divided by widths) should not exceed 20) and $N$ constraints on the stress, calculated at the left end of each segment (stresses should not exceed  $\bar{\sigma}=14000\; N/cm^2$).
-There is also also a constraint on the displacement at the tip, which should not exceed 2.5 cm.
-The load is $P = 50 000\; N$, the Young's modulus is $E=2\cdot 10^7  N/cm^2$.
+The number of segments $N$ can be chosen arbitrarily. The total length of the
+beam is 500~cm, the lengths of the segments are $l_i=500/N$~cm. There are $N$
+geometric constraints (the aspect ratios of the blocks, i.e. heights divided by
+widths, should not exceed 20) and $N$ constraints on the stress, calculated at
+the left end of each segment (stresses should not exceed
+$\bar{\sigma}=14\,000\;\text{N}/\text{cm}^2$). There is also a constraint on
+the displacement at the tip, which should not exceed 2.5~cm. The load is $P =
+50\,000$\;N; the Young's modulus is $E=2\cdot 10^7\;\text{N}/\text{cm}^2$.
 
-The deflection $y_i$ at the right hand of the $i$-th segment is given by the following recursive formulas:
+The deflection $y_i$ at the right end of the $i$th segment is given by the
+following recursive formulas:
 \begin{displaymath}
@@
 \end{displaymath}
-The moment of inertia of a segment is $I_i={b_i h_i^3}/{12}$ and the bending moment at the left end is $M_i=P[L+l_i- \sum_{j=1}^{i}  l_j ]$.
-The maximum bending stress in the segment $i$ is then given by the following formula:
+The moment of inertia of the $i$th segment is $I_i={b_i h_i^3}/{12}$, and the
+bending moment at its left end is $M_i=P[L+l_i- \sum_{j=1}^{i}  l_j ]$. The
+maximum bending stress in the $i$th segment is then given by the following
+formula:
 \begin{displaymath}
@@
 
-We are looking for a design with smallest volume $V = \sum_{i=1}^N b_i h_i l_i$.
-The widths $b_i$ vary from 1.0 to 10.0 cm and the heights $h_i$ from 5.0 to 100.0 cm.
-The optimization problem is formulated as follows:
+We should look for a design of smallest volume $V = \sum_{i=1}^N b_i h_i l_i$.
+The widths $b_i$ vary from 1.0 to 10.0~cm and the heights $h_i$ from 5.0 to
+100.0~cm. The optimization problem is formulated as follows:
 \begin{displaymath}
@@
     \min\limits_{ b,  h}V( b,  h) \\
-    s.t.\;1.0\le b_i \le 10.0, \\
+    \text{s.t.}\ 1.0\le b_i \le 10.0, \\
     5.0 \le h_i \le 100.0, \\
@@
 
-With $N=50$ segments (corresponding to 100 design variables) the SQP solution of this problem is $V = 63704.598\; cm^3$.
-The optimal values of design variables are given below (the first 50 entries are the widths and the last 50 entries are the heights of the segments):
+With $N=50$ segments (corresponding to 100 design variables), the SQP solution
+of the problem is $V = 63704.598\;\text{cm}^3$. The optimal values of the
+design variables are given below (the first 50 entries are the widths, and the
+last 50 entries are the heights of the segments):
 \setcounter{MaxMatrixCols}{11}
@@
 			  & 2.379,& 2.338,& 2.295,& 2.250,& 2.204,& 2.156,& 2.105,& 2.052,& 1.996,& 1.936, \\
-			  & 1.873,& 1.805,& 1.732,& 1.651,& 1.562,& 1.462,& 1.345,& 1.196,& 1.023,& \left. 1.000\right] 	
+			  & 1.873,& 1.805,& 1.732,& 1.651,& 1.562,& 1.462,& 1.345,& 1.196,& 1.023,& \left. 1.000\right], 	
 \end{matrix}
@@
 		& 47.582,&46.754,&45.897,&45.008,&44.082,&43.116,&42.105,&41.041,&39.919,&38.729  \\
-		& 37.462,&36.102,&34.630,&33.020,&31.240,&29.241,&26.910,&23.919,&20.465,&\left.14.639 \right]
+		& 37.462,&36.102,&34.630,&33.020,&31.240,&29.241,&26.910,&23.919,&20.465,&\left.14.639 \right].
 \end{matrix}
@@
 
-MAM obtained the solution $V = 63935.360\; cm^3$, using 2201 function evaluations (as compared to almost 10000 evaluations used by SQP).
-The number of points in a trust region used to build approximations was 200.
-The optimal values of design variables obtained by MAM are given below:
+By MAM, we obtained the solution $V = 63935.360\; \text{cm}^3$, using 2201
+function evaluations (as compared to almost 10\,000 evaluations used by SQP).
+The number of points in the trust region used to build the approximations was
+200. The optimal values of the design variables obtained by MAM are given
+below:
 \[
@@
 		& 2.368,&2.310,&2.307,&2.227,&2.176,&2.149,&2.106,&2.016,&2.007,&1.925 \\
-		& 1.864,&1.843,&1.758,&1.635,&1.582,&1.934,&1.332,&1.173,&1.026,&\left.2.419\right]
+		& 1.864,&1.843,&1.758,&1.635,&1.582,&1.934,&1.332,&1.173,&1.026,&\left.2.419\right],
 \end{matrix}
@@
 		& 47.373,&46.190,&46.140,&44.543,&43.525,&42.991,&42.119,&40.308,&40.141,&38.506  \\
-		& 37.268,&36.839,&35.153,&32.702,&31.615,&28.304,&26.642,&23.447,&20.536,& \left.9.417\right]
+		& 37.268,&36.839,&35.153,&32.702,&31.615,&28.304,&26.642,&23.447,&20.536,& \left.9.417\right].
 \end{matrix}
@@
 
-The solution obtained by MAM is very close to the reference solution obtained by SQP, except for the last design variable (the height of the last segment), which indicates that the problem is insensitive to this variable near the optimum, making it hard for the metamodels to capture this dependence.
-Both SQP and MAM solutions are, however, feasible and differ only slightly in the value of the objective function.
+The solution obtained by MAM is very close to the reference solution obtained
+by SQP, except for the last design variable (the height of the last segment),
+which indicates that the problem is insensitive to this variable near the
+optimum, making it hard for metamodels to capture this dependence. Both SQP
+and MAM solutions are, however, feasible and differ only slightly in the value
+of the objective function.
 
-Next, let us compare the worktime of the sequential and parallel algorithms when solving the large scale problems. 
-The dimensionality of the problem being solved $N$ was varied from 100 up to 1000 design variables that corresponds to 50 to 500 segments of the cantilevered beam.
-The number of points in a trust region used to build the approximations was $2 N$.
-Both algorithms was run on a single node of the cluster (the specifications of the node are listed below), the parallel algorithm employed all 16 processors cores available.
-Since the time of computing of the objective
-function and constraints in the test problem being solved was negligible, these ones were computed on the same node.
+Next, let us compare the work time of the sequential and parallel algorithms
+when solving large-scale problems. The dimensionality $N$ of the problem being
+solved was varied from 100 up to 1000 design variables, which corresponds to a
+variation of the number of segments of the cantilevered beam from 50 to 500.
+The number of points in the trust region used to build the approximations was
+$2 N$. Both algorithms were run on a single node of the cluster (the
+specifications of the node are listed below), the parallel algorithm employed
+all 16 processors cores available. Since the time of computing the objective
+function and the constraints in the test problem was negligible, these were
+computed on the same node.
 
@@
 
-Table 1 reflects the worktime (in seconds) of the sequential algorithm and of the parallel one subject to the number of variables.
-In Table 2, the number of MAM iterations $I_{MAM}$ as well as the number of the function evaluations $F_{MAM}$ required for solving the problem are presented.
-For comparison, the number of the function value computations, which would be required to solve the initial (non-approximated) problem by SQP method $F_{SQP}$ is given also.
-In all conducted experiments, the objective function values in the sequential and parallel versions of the algorithm were the same (up to the accuracy of the computational error) and differed from the solution obtained by SQP negligibly.
-The reduction of the number of the function evaluations required to solve the problem using MAM as compared to the use of SQP was demonstrated
-visibly.
+Table 1 reflects the work time (in seconds) of the sequential algorithm and
+that of the parallel one subject to the number of variables. The number of MAM
+iterations $I_{MAM}$ as well as the number of function evaluations $F_{MAM}$
+required for solving the problem are presented in Table 2. For comparison
+purposes, the number of function value computations $F_{SQP}$ that would be
+required to solve the initial (non-approximated) problem by the SQP method is
+also given. In all conducted experiments, the objective function values in the
+sequential and parallel versions of the algorithm were the same (up to
+computational errors) and negligibly differed from the solution obtained by
+SQP. The reduction of the number of the function evaluations required to solve
+the problem using MAM as compared to the use of SQP was demonstrated visibly.
 
-Computational experiments were carried out on a high-performance cluster of Lobachevsky State University of Nizhny Novgorod.
-The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.
-The CPU had 8 cores, i.e.
-a total of 16 physical cores were available in the node.
-MS Visual Studio 15 and Intel Fortran Compiler were used to implement the algorithm.
+The computational experiments were carried out on a high-performance cluster at
+Lobachevsky State University of Nizhny Novgorod. A cluster node includes two
+Intel Sandy Bridge E5-2660 2.2~GHz CPUs and 64~Gb~RAM. Each CPU has 8 cores,
+i.e. a total of 16 physical cores were available at the node. MS Visual Studio
+15 and Intel Fortran Compiler were used to implement the algorithm.
 
@@
 
-Recent developments in the Multipoint Approximation Method (MAM) made it capable of solving large scale industrial optimization problems.
-The fact that MAM solves the initial problem by using the approximations of the objective function and constraints is the primary distinctive feature of the method.
-Within the framework of the present study, the parallel version of MAM oriented onto the reduction of the worktime of the optimization algorithm (assuming that computing the values of the objective function and constraints has been parallelized already) has been developed.
-The performed experiments have demonstrated an acceptable speedup when solving the large scale problems employing 16 cores on a single cluster node.
-The performance was demonstrated on a benchmark example of structural
-optimization known as the scalable cantilevered beam.
+Recent developments in the Multipoint Approximation Method (MAM) made it
+capable of solving large-scale industrial optimization problems. The fact that
+MAM solves the initial problem by using approximations of the objective
+function and constraints is the primary distinctive feature of the method.
+Within the framework of the present study, we developed a parallel version of
+MAM oriented to the reduction of the work time of the optimization algorithm
+(assuming that the computation of the values of the objective function and
+constraints has already been parallelized). The experiments performed have
+demonstrated an acceptable speedup when solving large-scale problems employing
+16~cores on a single cluster node. The performance was demonstrated on a
+benchmark example of structural optimization known as the scalable cantilevered
+beam.
 
@@
 \newblock Simulation approach to structural optimization.
-\newblock Structural Optimization \textbf{1}, 37--46 (1989)
+\newblock Structural Optimization \textbf{1}, 37--46 (1989).
 
@@
 \newblock In: Sydow, A. (ed.) Proceedings of the 4th international symposium on Systems analysis and simulation,
-pp. 207--212, Elsevier,  (1992)
+pp.\,207--212, Elsevier,  (1992).
 
@@
 \newblock Multiparameter structural optimization using FEM and multipoint explicit approximations.
-\newblock Structural Optimization \textbf{6}, 7--14 (1993)
+\newblock Structural Optimization \textbf{6}, 7--14 (1993).
 
@@
 \newblock Large scale optimization of transonic axial compressor rotor blades.
-\newblock In:  49th AIAA/ASME/ASCE/AHS/ASC Structures, Structural Dynamics, and Materials Conference, AIAA, art. no. 2008-2056 (2008)
+\newblock In:  49th AIAA/ASME/ASCE/AHS/ASC Structures, Structural Dynamics, and Materials Conference, AIAA, art. no. 2008-2056
+(2008).
 
@@
 \newblock Design optimization of aircraft engine components.
-\newblock In: Proc. of 7th ASMO UK / ISSMO Conf. on Engineering Design Optimization. Process and Product Improvement (2008)
+\newblock In: Proc. of 7th ASMO UK / ISSMO Conf. on Engineering Design Optimization. Process and Product Improvement
+(2008).
 
@@
 \newblock Multidisciplinary optimization of turbomachinery based on metamodel built by genetic programming.
-\newblock In: 13th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference (2010)
+\newblock In: 13th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference
+(2010).
 
@@
 \newblock In: McCarthy, J.M. (ed.) Proc. of ASME Design Engineering Technical Conferences and Computers
-  in Engineering Conf., 1--12, ASME, Irvine CA (1996)
+  in Engineering Conf., 1--12, ASME, Irvine CA (1996).
 
@@
 \newblock Refinements in the multi-point approximation method to reduce the effects of noisy responses.
-\newblock In: 6th AIAA/NASA/ISSMO Symp. on Multidisciplinary Analysis and Optimization,  pp. 941--951, AIAA (1996)
+\newblock In: 6th AIAA/NASA/ISSMO Symp. on Multidisciplinary Analysis and Optimization,  pp.\,941--951, AIAA
+(1996).
 
@@
 \newblock The use of simplified numerical models as mid-range approximations.
-\newblock In: 6th AIAA/NASA/ISSMO Symp. on Multidisciplinary Analysis and Optimization, pp. 952--958 (1996).
+\newblock In: 6th AIAA/NASA/ISSMO Symp. on Multidisciplinary Analysis and Optimization, pp.\,952--958 (1996).
 
@@
 \newblock In: Parmee, I.C. (ed.) Adaptive Computing in Design and Manufacture, 3rd Int. Conf.,
-   pp. 21--24, PEDC, Dartington (1998)
+   pp.\,21--24, PEDC, Dartington (1998).
 
@@
 \newblock Mid-range metamodel assembly building based on linear regression for large scale optimization problems.
-\newblock Structural and Multidisciplinary Optimization \textbf{45}(4), 515--527 (2012)
+\newblock Structural and Multidisciplinary Optimization \textbf{45}(4), 515--527
+(2012).
 
@@
 \newblock Surfaces generated by moving least squares methods.
-\newblock Mathematics of Computation \textbf{87}, 141--158 (1981)
+\newblock Mathematics of Computation \textbf{87}, 141--158 (1981).
 
@@
 \newblock An interpolation method for an irregular net of nodes.
-\newblock Int. J. Num. Meth. \textbf{20}, 1599--1612 (1984)
+\newblock Int. J. Num. Meth. \textbf{20}, 1599--1612 (1984).
 
@@
 \newblock Moving least squares method for reliability-based design optimization.
-\newblock In: 4th World Congress of Structural and Multidisciplinary Optimization, Dalian, China (2001)
+\newblock In: 4th World Congress of Structural and Multidisciplinary Optimization, Dalian, China
+(2001).
 
@@
 \newblock Design optimization and stochastic analysis based on the moving least squares method.
-\newblock In: Herskovits, J., Mazorche, S., Canelas, A. (eds.) 6th World Congress of Structural and Multidisciplinary Optimization, art. no. 9412 (2005)
+\newblock In: Herskovits, J., Mazorche, S., Canelas, A. (eds.) 6th World Congress of Structural and Multidisciplinary Optimization, art. no. 9412
+(2005).
 
@@
 \newblock Recognition of design variable inter-dependencies using cross-validated moving least-squares method.
-\newblock In: Proceedings of the 51st AIAA/ASME/ASCE/AHS/ASC Structures, Structural Dynamics, and Materials Conference, AIAA, art. no. 2010-2985 (2010)
+\newblock In: Proceedings of the 51st AIAA/ASME/ASCE/AHS/ASC Structures, Structural Dynamics, and Materials Conference, AIAA, art. no. 2010-2985
+(2010).
 
@@
 \newblock  Mid-range approximations in sub-spaces for MDO problems with disparate discipline attributes.
-\newblock In: 15th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference, AIAA, art. no. 2014-2437 (2014)
+\newblock In: 15th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference, AIAA, art. no. 2014-2437
+(2014).
 
@@
 \newblock New developments in structural optimization using adaptive mesh refinement and multi-point approximations.
-\newblock Engineering Optimization \textbf{29}, 217--234 (1997)
+\newblock Engineering Optimization \textbf{29}, 217--234 (1997).
 
@@
 \newblock Adaptive and parallel capabilities in the multipoint approximation  method.
-\newblock In: 12th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference, AIAA, art. no. 2008-5803 (2008)
+\newblock In: 12th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference, AIAA, art. no. 2008-5803
+(2008).
 
@@
 \newblock Multidiscipline Design Optimization.
-\newblock Vanderplaats Research \& Development, Inc., Colorado Springs, (2001)
+\newblock Vanderplaats Research \& Development, Inc., Colorado Springs, (2001).
 

### 2018-05-10 2018/YSC/ysc_abstract.tex
YSC: add abstract

% easychair.tex,v 3.1 2011/12/30
%
% Select appropriate paper format in your document class as
% instructed by your conference organizers. Only withtimes
% and notimes can be used in proceedings created by EasyChair
%
% The available formats are 'letterpaper' and 'a4paper' with
% the former being the default if omitted as in the example
% below.
%
\documentclass[procedia]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

% This provides the \BibTeX macro
\usepackage{doc}
\usepackage{makeidx}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% If you plan on including some algorithm specification, we recommend
% the below package. Read more details on the custom options of the
% package documentation.
%
% \usepackage{algorithm2e}

\def\procediaConference{99th Conference on Topics of
  Superb Significance (COOL 2014)}

%\makeindex

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Comparison of dimensionality reduction schemes for parallel global optimization
algorithms}

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair

\titlerunning{Dimension reduction schemes in parallel GO}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes into the list
% defined using \institute
%
\author {
Vladislav Sovrasov \and
Ilya Lebedev
}

% Institutes for affiliations are also joined by \and,
\institute  {
  Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
  \email{sovrasov.vlad@gmail.com}\\
  \email{ilya.lebedev@itmm.unn.ru}\\
}
%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{Sovrasov V., Lebedev I.}

\begin{document}

\maketitle

\keywords{Global optimization, Dimension reduction, Parallel algorithms,
Multidimensional multiextremal optimization, Global search algorithms}

\begin{abstract}
This work considers a parallel algorithms for solving multi-extremal optimization problems.
Algorithms are developed within the framework of the information-statistical approach and
implemented in a parallel solver Globalizer. The optimization problem is solved by reducing
the multidimensional problem to a set of joint one-dimensional problems that are solved in
parallel. Five types of Peano-type space-filling curves are employed to reduce dimension. The
results of computational experiments carried out on several hundred test problems are discussed.
\end{abstract}

%\setcounter{tocdepth}{2}
%{\small
%\tableofcontents}

%\section{To mention}
%
%Processing in EasyChair - number of pages.
%
%Examples of how EasyChair processes papers. Caveats (replacement of EC
%class, errors).

%------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}

In the present paper, the parallel algorithms for solving the multiextremal optimization problems
are considered. In the multiextremal problems, the opportunity of reliable estimate of the global
optimum is based principally on the availability of some information on the function known
{\textit a priori} allowing relating the probable values of the optimized function to the known
values at the points of performed trials. Very often, such an information on the problem being
solved is represented in the form of suggestion that the objective function $\varphi(y)$ satisfies
Lipschitz condition with the constant $L$ not known a priori (see, for example,
\cite{Jones,Gablonsky,Evtushenko}). At that, the objective function could be defined by a
program code i. e. could represent a ``black-box''-function. Such problems are presented in the
applications widely (problems of optimal design of objects and technological processes in
various fields of technology, problems of model fitting according to observed data in scientific
research, etc.).

Many methods destined to solving the problems of the class specified above reduce the solving
of a multidimensional problem to solving the one-dimensional subproblems implicitly (see, for
example, the methods of diagonal partitions \cite{Sergeyev2006,SergeyevKvasov2015} or
simplicial partitions \cite{Zilinskas2008,Zilinskas2014}). In the present work, we will use the
approach developed in Lobachevsky State University of Nizhni Novgorod based on the idea of
the dimensionality reduction with the use of Peano space-filling curves $y(x)$ mapping the
interval $[0,1]$ of the real axis onto an $n$-dimensional cube continuously and unambiguously.

Several methods of constructing the evolvents approximating the theoretical Peano curve have
been proposed in \cite{strongin1978,Strongin1992,Goryachih2017,Gergel2009}. These
methods were implemented in the Globalizer software system \cite{globalizerSystem}. The goal
of the present study was comparing the properties of the evolvents and the selecting the most
suitable ones for the use in the parallel global optimization algorithms.

%------------------------------------------------------------------------------
% Index
%\printindex
\begin{thebibliography}{107}

\bibitem{strongin1978}% (MR509033)
\newblock R. G. Strongin,
\newblock \emph{Numerical Methods in Multi-Extremal Problems (Information-Statistical Algorithms)},
\newblock Moscow: Nauka (1978) (In Russian)

\bibitem{Strongin1992}% (MR1263606) [10.1007/BF00122428]
\newblock R. G. Strongin,
\newblock \emph{\emph{Algorithms for multi-extremal mathematical programming problems
employing a set of joint space-filling curves}},
\newblock \emph{J. Glob. Optim.}, \textbf{2}, 357--378 (1992)

\bibitem{Goryachih2017}
\newblock Goryachih, A.
\newblock \emph{A class of smooth modification of space-filling curves for global optimization
problems}
\newblock Springer Proceedings in Mathematics and Statistics, \textbf{197}, pp. 57--65 (2017)

\bibitem{Gergel2009}
Strongin, R.G., Gergel, V.P., Barkalov, K.A.: \emph{Parallel methods for global optimization problem
solving.} Journal of instrument engineering. \textbf{52}, 25--33 (2009) (In Russian)

\bibitem{globalizerSystem}
Gergel V.P., Barkalov K.A., and Sysoyev A.V: Globalizer: \emph{A novel supercomputer software
system for solving time-consuming global optimization problems.} Numerical Algebra, Control
\& Optimization \textbf{8(1)}, 47--62 (2018)

\bibitem{Sergeyev2006}
Sergeyev, Ya.D., Kvasov, D.E.: \emph{Global search based on efficient diagonal partitions and a set
of Lipschitz constants}. SIAM J. Optim \textbf{16(3)}, 910--937 (2006)

\bibitem{SergeyevKvasov2015}
Sergeyev, Y.D., Kvasov, D.E.: \emph{A deterministic global optimization using smooth diagonal
auxiliary functions.} Communications in Nonlinear Science and Numerical Simulation. \textbf{21(1-3)},
99--111 (2015)

\bibitem{Zilinskas2008}
\v Zilinskas, J.: \emph{Branch and bound with simplicial partitions for global optimization.} Math.
Model. Anal. \textbf{13(1)}, 145--159 (2008)

\bibitem{Zilinskas2014}
Paulavi\v cius, R., \v Zilinskas, J.: \emph{Simplicial Lipschitz optimization without the Lipschitz
constant.} J. Glob. Optim. \textbf{59(1)}, 23-40 (2014)

\bibitem{Jones}
Jones, D.R.: The direct global optimization algorithm. In: Floudas, C.A., Pardalos, P.M.
(eds.) \emph{The Encyclopedia of Optimization}, 2nd edn., pp. 725--735. Springer, Heidelberg
(2009)

\bibitem{Gablonsky}
Gablonsky, J.M., Kelley, C.T.: \emph{A locally-biased form of the DIRECT algorithm.} J. Glob.
Optim., \textbf{21(1)}, 27--37 (2001)

\bibitem{Evtushenko}
Evtushenko, Y., Posypkin, M.: \emph{A deterministic approach to global box-constrained
optimization.} Optim. Lett. \textbf{7(4)}, 819--829 (2013)

\end{thebibliography}

%------------------------------------------------------------------------------
\end{document}

% EOF

### 2018-05-10 2018/Journal of Supercomputing/JoS_Paper.tex
Major corrections of the language

@@
 \begin{abstract}
-In the present work, a parallel algorithm for solving the multiextremal problems with non-convex constrains is considered.
-Within the framework of approach considered, the solving of initial multidimensional problem is reduced to the parallel solving of a set of linked one-dimensional problems.
-The parallelization scheme is featured by the absence of a single control process since the solution of any one-dimensional problem is a solution of the initial multidimensional one.
-The non-convex constraints are accounted for by a novel index scheme not utilizing the ideas of penalty functions.
-The efficiency of the parallel algorithm was confirmed experimentally by the numerical solving of several hundred multidimensional multiextremal problems with non-convex constrains generated randomly.
+This paper considers a parallel algorithm for solving multiextremal problems with non-convex constrains.
+Within the framework of the approach proposed here, the solution of the initial multidimensional problem is reduced solving a set of linked one-dimensional problems in parallel.
+The parallelization scheme does not require a single control process since the solution of any one-dimensional problem is a solution of the initial multidimensional one.
+The non-convex constraints are accounted for by a novel index scheme without utilizing the ideas of penalty functions.
+The efficiency of the parallel algorithm was confirmed experimentally by solving numerically of several hundred multidimensional multiextremal problems with non-convex constrains generated randomly.
 \keywords{Global optimization \and Constrained problems \and Non-convex constraints \and Dimensionality reduction \and Parallel algorithms}
@@
 
-This work considers parallel methods for solving global optimization problems with non-convex constraints.
+The aim of this paper is to examine parallel methods for solving global optimization problems with non-convex constraints.
 The objective function and constraints are assumed to satisfy the Lipschitz condition with a priori unknown Lipschitz constants.
 The analytical form of the problem’s functions may also be unknown, i.e.
-they can be set by an algorithm computing their values at various points within the search domain (so-called ``black-box'' functions).
-Moreover, it is supposed that even a single computation of the function value can be time-consuming, as in the applied problems it requires performing numerical simulation.
+they can be set by an algorithm computing their values at various points within the search domain (the so-called ``black-box'' functions).
+It is also supposed that even a single computation of the function value can be time-consuming, since it requires performing numerical simulation in applied problems.
 %These assumptions are typical for many approaches to building parallel algorithms for unconstrained global optimization \cite{Jones2001,Paulavicius2011,Evtushenko2013}.
 
-The numerical algorithms for solving the problems of the specified class are featured by an exponential growth of the computational costs with increasing dimensionality (so called ``curse of dimensionality'').
-The use of supercomputers expands the field of application of the global optimization methods and, at the same time, sets a problem of efficient parallelization of the search process.
+The numerical algorithms for solving the problems of the above-mentioned class are noted for an exponential growth of the computational costs with increasing dimensionality (the so-called ``curse of dimensionality'').
+The use of supercomputers expands the field of application of the global optimization methods and, however, at the same time, it gives rise to the problem of efficient parallelization of the search process.
 
-At present, the parallel algorithms, which utilize the heuristic approaches and based on the idea of random search in any way are quite popular (see \cite{RefFerreiro,RefZhu,Korosec,Guerrero}).
-However, because of the probabilistic nature, the algorithms of such type guarantee the convergence to the global optimum solution in the probabilistic meaning only, that distinguishes these ones unfavorably from the deterministic parallel global optimization algorithms \cite{Jones2001,Paulavicius2011,Evtushenko2013}.
-The efficient deterministic algorithms are oriented onto the solving of the unconstrained optimization problems as a rule.
-The problems with constraints are solved using the penalty function method usually having a number of disadvantages.
-One of these is the impossibility of its application in the problems with the \textit{partially defined functions} (i.e.
-when violating any constraint, the values of the rest functions of the problem remain undefined).
+Parallel algorithms utilizing heuristic approaches and based in one way or another on the idea of random search are at present quite popular (see \cite{RefFerreiro,RefZhu,Korosec,Guerrero}).
+However, because of their probabilistic nature, the algorithms of such type guarantee the convergence to the global optimum solution in the probabilistic sense only, which makes them less attractive than deterministic parallel global optimization algorithms \cite{Jones2001,Paulavicius2011,Evtushenko2013}.
+As a rule, efficient deterministic algorithms are intended to solve unconstrained optimization problems.
+The problems with constraints are solved using the penalty function method that has a number of disadvantages.
+In particular, this method cannot be applied to the problems with the \textit{partially defined functions} (i.e.
+when any constraint is violated, the values of all the other functions of the problem remain undefined).
 
-As the examples, one can mention the optimal control problems, described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2011,Balandin2017}.
+Among the examples of such problems are the optimal control problems described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2011,Balandin2017}.
 It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e.
-every eigenvalue of $A$ has strictly negative real part.
-Otherwise the value of the criteria is indeterminate.
+every eigenvalue of $A$ has a strictly negative real part.
+Otherwise, the value of the criteria is indeterminate.
 
-In the present work, a novel approach to the minimization of the multiextremal functions with non-convex constraints developed in \cite{Strongin2000,Sergeyev2001,Barkalov2002} and referred to as the index method  of accounting for the constraints was applied.
-The approach is based on a separate accounting for each constraint of the problem and is not related to the use of the penalty functions.
-The solving of initial multidimensional problem is reduced (using the Peano-type space-filling curves) to solving an equivalent set of the one-dimensional problems, which can be solved in parallel in such a way that each of these processes shares the information obtained by the other processes.
-The propose scheme was implemented using Lobachevsky supercomputer with the use of the distributed memory as well as the shared one.
+In this study, we used a novel approach to the minimization of multiextremal functions with non-convex constraints developed in \cite{Strongin2000,Sergeyev2001,Barkalov2002}.
+The approach referred to as the index method of accounting for the constraints is based on accounting for each constraint of the problem separately and is not related to the use of the penalty functions.
+The solution of the initial multidimensional problem is reduced (by using the Peano-type space-filling curves) to solving an equivalent set of one-dimensional problems, which can be solved in parallel in such a way that each of these processes shares the information obtained by the other processes.
+The proposed scheme was implemented on the Lobachevsky supercomputer with the use of the distributed memory as well as the shared one.
 
@@
 \end{equation}
-The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy Lipschitz condition 
+The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
 \[
@@
 \]
-with a priory unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal.
-It is assumed that functions $g_i(y)$ are defined and computable only at the points $y \in D$ satisfying the conditions
+with a priori unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal.
+It is assumed that the functions $g_i(y)$ are defined and computable only at the points $y \in D$ satisfying the conditions
 \begin{equation}\label{g_k}
@@
 
-Employing the continuous single-valued Peano curve $y(x)$ mapping the unit interval $[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}) it is possible to find the minimum in (\ref{problem}) by solving the one-dimensional problem
+By employing the continuous single-valued Peano curve $y(x)$ mapping the unit interval $[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}), it is possible to find the minimum in (\ref{problem}) by solving the one-dimensional problem
 \[
@@
 
-These conditions allows to introduce a classification of the points $x \in [0,1]$ according to the number $\nu (x)$ of the constraints computed at this point.
+These conditions allows us to introduce a classification of the points $x \in [0,1]$ according to the number $\nu (x)$ of the constraints computed at this point.
 The \textit{index} $\nu(x)$ can also be defined by the conditions
@@
 
-The considered dimensionality reduction scheme juxtaposes to a multidimensional problem with lipschitzian functions a one-dimensional problem, where the corresponding functions satisfy uniform H{\"o}lder condition (see \cite{Sergeyev2013}), i.e.,
+In the dimensionality reduction scheme considered here, a multidimensional problem with Lipschitzian functions is juxtaposed with a one-dimensional problem, where the corresponding functions satisfy uniform H{\"o}lder condition (see \cite{Sergeyev2013}), i.e.,
 \[
@@
 \]
-Here $N$ is the dimensionality of the initial multidimensional problem and the coefficients $H_i$ are related with Lipschitz constant $L_i$ of the initial problem as $H_i \leq 2L_i \sqrt{N+3}$.
+Here, $N$ is the dimensionality of the initial multidimensional problem and the coefficients $H_i$ are related to the Lipschitz constant $L_i$ of the initial problem as $H_i \leq 2L_i \sqrt{N+3}$.
 
-Thus, \textit{a trial} at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm will consist in the following sequence of operations:
+Thus, \textit{a trial} at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm will consist of the following sequence of operations:
 \begin{itemize}
@@
 
-Step 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the lower indices in increasing order of the coordinate values, i.e.
+Step 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the lower indices in the increasing order of the coordinate values, i.e.
 \[
@@
 
-Rule 2. For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} $R(i)$ using some formulae.
+Rule 2. For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} $R(i)$ using some formulae \cite{Barkalov2002}.
 
@@
 
-Rule 4. Execute the next trial in the internal point of the interval $(x_{t-1},x_t)$, i.e. $x^{k+1} \in (x_{t-1},x_t)$.
+Rule 4. Execute the next trial at the internal point of the interval $(x_{t-1},x_t)$, i.e. $x^{k+1} \in (x_{t-1},x_t)$.
 
@@
 
-Detailed description of this algorithm and the corresponding theory of convergence are presented in \cite{Strongin2000,Sergeyev2001,Barkalov2002}.
+A detailed description of this algorithm and the corresponding theory of convergence are presented in \cite{Strongin2000,Sergeyev2001,Barkalov2002}.
 
-\section{Parallel Index Algorithm with the Set of Evolvents}
+\section{Parallel Index Algorithm with a Set of Evolvents}
 \label{sec:3}
 
-The reduction of the multidimensional problems to the one-dimensional ones using evolvents has such important properties as the continuity and preservation of boundedness of function divided differences.
-However, a partial loss of information on the nearness of the points in the multidimensional space takes place since a point $x \in [0,1]$ has only the left and the right neighbors while the corresponding point $y(x) \in R^N$ has the neighbors in $2N$ directions.
-As a result, when using the mappings like Peano curve the images $y', y''$, which are close to each other in the $N$-dimensional space can correspond to the preimages $x', x''$, which can be far away from each other in the interval $[0,1]$.
-This property results in the excess computations since several limit points $x', x''$ of the trial sequence generated by the index method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional space.
+The reduction of the multidimensional problems to the one-dimensional ones using evolvents has such important properties as the continuity and preservation of boundedness of function's divided differences.
+However, a partial loss of information on the nearness of the points in the multidimensional space occurs in this case since a point $x \in [0,1]$ has only the left and the right neighbors while the corresponding point $y(x) \in R^N$ has the neighbors in $2N$ directions.
+As a result, when using the mappings like a Peano curve, the images $y', y''$, which are close to each other in the $N$-dimensional space can correspond to the preimages $x', x''$, which can be far away from each other in the interval $[0,1]$.
+This property results in excess computations since several limit points $x', x''$ of the trial sequence generated by the index method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional space.
 
-One of the possible ways to overcome this disadvantage consists in using the multiple mapping $Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of single evolvent $y(x)$.
-To construct the set $Y^S(x)$ different approaches can be used.
+One of the possible ways to overcome this disadvantage consists in using the multiple mapping $Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of the single evolvent $y(x)$.
+To construct the set $Y^S(x)$, different approaches can be used.
 For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^i(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$.
@@
 
-Using the multiple mapping allows solving initial problem (\ref{problem}) by parallel solving the problems
+With the use of multiple mapping, it is possible to solve the initial problem (\ref{problem}) (1) by solving in parallel the problems
 \[
@@
 Each one-dimensional problem is solved on a separate processor/node.
-The trial results at the point $x^k$ obtained for the problem being solved by particular processor are interpreted as the results of the trials in the rest problems (at the corresponding points $x^{k_1},...,x^{k_S}$).
-In this approach, a trial at the point $x^k \in [0,1]$ executed in the framework of the $s$-th problem, consists in the following sequence of operations.
+The trial results at the point $x^k$ obtained for the problem being solved by particular processor are interpreted as the results of the trials in the remaining problems (at the corresponding points $x^{k_1},...,x^{k_S}$).
+In this approach, a trial at the point $x^k \in [0,1]$ executed in the framework of the $s$-th problem, consists of the following sequence of operations.
 
@@
 	\]
-The occurrence of the first violation of the constraint terminates the trial.
-In the case when the point $y^k$ is a feasible one the trial includes the computation of the values of all problem functions.
+The trial terminates at the point as soon as the first violation of any constraint occurs.
+In the case when $y^k$ is a feasible one, the trial includes the computation of the values of all problem functions.
 In this situation, the index is set to $\nu=m+1$.
@@
 
-\item Determine the preimages $x^{k_s} \in [0,1]$, $1 \leq s \leq S$, of the point $y^k$ and interpret the trial executed at the point $y^k \in D$ as the execution of the trials in the $S$ points  $x^{k_1},...,x^{k_S}$ with the same results
+\item Determine the preimages $x^{k_s} \in [0,1]$, $1 \leq s \leq S$, of the point $y^k$ and interpret the trial executed at the point $y^k \in D$ as the execution of the trials at the $S$ points  $x^{k_1},...,x^{k_S}$ with the same results
 \[
@@
 
-The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the sequential algorithm (except the method of the trial execution).
-For the organization of the interactions among the processors, the queues are created on each processor, where the processors store the information on the executed iterations in the form of the tuples: the processor number $s$, the trial point $x^{k_s}$, the index $\nu(x^{k_s})$, and the value $g_\nu(y^s (x^{k_s} ) )$.
+The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the sequential algorithm (with the exception the method of the trial execution).
+To organize the interactions between the processors, the queues are created on each processor, where the processors store the information on the executed iterations in the form of tuples: the processor number $s$, the trial point $x^{k_s}$, the index $\nu(x^{k_s})$, and the value $g_\nu(y^s (x^{k_s} ) )$.
 Moreover, the index of the blocked point is assumed to be equal to $-1$; the function value at this point is undefined.
@@
 A separate MPI-process is created for each of $S$ one-dimensional problems being solved, one process per one processor employed.
-Each process use $p$ threads for solving one-dimensional problem, one thread per an accessible core.
+Each process uses $p$ threads for solving a one-dimensional problem, one thread per an accessible core.
 
@@
 
-A well-known approach to the investigation and comparing of the multiextremal optimization algorithms is based on testing these methods by solving a set of problems, chosen randomly from some specially designed class.
+A well-known approach to investigating and comparing the multiextremal optimization algorithms is based on testing these methods by solving a set of problems, chosen randomly from some specially designed class.
 
@@
 
-In this study we will use GKLS generator of the functions of arbitrary dimensionality proposed in \cite{Gaviano2003} to produce the constrained problems.
-The scheme that allows to form the constrained global optimization problems is proposed in \cite{Gergel2017}.
-In the previous investigations, the index method has been confirmed experimentally to be not inferior to well-known analogues.
-The comparing of the method to well known DIRECT one \cite{Jones2001} in solving the unconstrained optimization problems has been performed in \cite{Barkalov2015,Barkalov2016}.
-In the present study, an experimental investigation of the speedup, which is obtained by the use of the index method in combination with the two-level parallelization scheme from \cite{Sidorov2015}.
+In this study, we will use a GKLS generator of the functions of arbitrary dimensionality proposed in \cite{Gaviano2003} to produce constrained problems.
+The scheme for generating constrained global optimization problems is proposed in \cite{Gergel2017}.
+Previous investigations have confirmed experimentally that the index method is not inferior to well-known analogues.
+A comparison of the method with the well known DIRECT method \cite{Jones2001} in solving unconstrained optimization problems has been performed in \cite{Barkalov2015,Barkalov2016}.
+In the present study, an experimental investigation is carried out of the speedup, which is obtained by the use of the index method in combination with the two-level parallelization scheme from \cite{Sidorov2015}.
 
@@
 The number of the used cluster nodes $S$ and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at each node have been varied.
-The problem was considered to be solved, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^*\right\|\leq \delta$.
-The size of the vicinity was selected as $\delta=0.03\left\|b-a\right\|$, where $a$ and $b$ are boundaries of the search domain.
-For the purpose of simulation of the computational complexity inherent to applied problems of optimization, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (series summation of $80$ thousand elements).
+The problem was considered to be solved, if the algorithm generated a trial point $y^k$ in the $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^*\right\|\leq \delta$.
+The size of the vicinity was selected as $\delta=0.03\left\|b-a\right\|$, where $a$ and $b$ are the boundaries of the search domain.
+For the purpose of simulation of the computational complexity inherent to applied optimization problems, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (a series summation of $80$ thousand elements).
 
@@
 The CPU had 8 cores, i.e.
-16 cores were available in each node.
+each node had a total of 16 cores.
 
-The average time and number of iterations, which were required to solve the problems of the series at various parallelization parameters are reflected in Tables 1 and 2.
-Here \textit{Node(core)} are the numbers of employed nodes and cores per a node, correspondingly.
+The average time and the number of iterations, which were required to solve the problems of the series at various parallelization parameters are given in Tables 1 and 2.
+Here, \textit{Node(core)} are the numbers of employed  nodes and cores per node, respectively.
 
@@
 
-The results demonstrate the presence of the speedup when using shared memory at a node (performing several trials within a problem in parallel) as well as the distributed memory (parallel solving of several subproblems at different nodes).
-At that, the highest time speedup was 10 (when using 64 cored on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster nodes).
+The results demonstrate the speedup achieved when using the shared memory at a node (performing several trials within a problem in parallel) as well as the distributed memory (solving in parallel several subproblems at different nodes).
+In this case, the highest time speedup was 10 (when using 64 cores on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster nodes).
 The difference in the speedups in time and in the number of iterations can be explained by the effect of the overheads of the data transmission between the processes.
-Note that when solving the applied optimization problems, the computing of the problem function values even in one point is a computation costly operation.
-The data transfer overheads will not affect the total computational costs predominately in this case, and the time speedup will not differ from the iteration one so strongly.
+Note that when solving applied optimization problems, the computation of the problem function values even in one point is a computationally costly operation.
+In this case, the data transfer overheads will not affect the total computational costs predominately in this case, and the difference between the time speedup and the iteration speedup will not be so great.
 %Описание инфраструктуры
@@
 
-The parallel index method for solving constrained global optimization problems considered in the present work:
+The parallel index method for solving constrained global optimization problems considered in this paper offers the following possibilities:
 \begin{itemize}
-	\item allows solving the initial problem directly, without the use of the penalty functions (thus, the issues of selection the penalty coefficient and of solving a series of unconstrained problems with different penalty coefficients are eliminated);
-	\item allows solving the problems, which the values of the problem function are not defined everywhere (for example, the objective function values are undefined out of the feasible domains of the problem constraints);
-	\item allows using the two-level parallelization scheme with the shared and distributed memory proposed earlier for the unconstrained optimization methods.
+	\item to solve the initial problem directly, without the use of the penalty functions (thus eliminating the need to select the penalty coefficient and to solve a series of unconstrained problems with different penalty coefficients);
+	\item to solve the problems where the values of the problem function are not defined everywhere (for example, the objective function values are undefined outside the feasible domains);
+	\item to use the two-level parallelization scheme with the shared and distributed memory proposed earlier for the unconstrained optimization methods.
 \end{itemize}
 
-The parallel algorithm has demonstrated speedup with respect to the number of processors/cores employed.
-This was confirmed by the results of the numerical solving of several hundred test problems using 128 cores of Lobachevsky supercomputer.
+The parallel algorithm has demonstrated a speedup with respect to the number of processors/cores employed.
+This was confirmed by the results of solving numerically of several hundred test problems using 128 cores of the Lobachevsky supercomputer.
 %The direction of further research is the generalization of the considered parallel algorithm for solving multicriteria problems.

### 2018-05-11 2018/Journal of Supercomputing/JoS_Paper.tex
Before final cheking

@@
 This paper considers a parallel algorithm for solving multiextremal problems with non-convex constrains.
-Within the framework of the approach proposed here, the solution of the initial multidimensional problem is reduced solving a set of linked one-dimensional problems in parallel.
+Within the framework of the approach proposed here, the solution of the initial multidimensional problem is reduced to solving a set of linked one-dimensional problems in parallel.
 The parallelization scheme does not require a single control process since the solution of any one-dimensional problem is a solution of the initial multidimensional one.
@@
 Parallel algorithms utilizing heuristic approaches and based in one way or another on the idea of random search are at present quite popular (see \cite{RefFerreiro,RefZhu,Korosec,Guerrero}).
-However, because of their probabilistic nature, the algorithms of such type guarantee the convergence to the global optimum solution in the probabilistic sense only, which makes them less attractive than deterministic parallel global optimization algorithms \cite{Jones2001,Paulavicius2011,Evtushenko2013}.
+However, because of their probabilistic nature, the algorithms of such type guarantee the convergence to the global optimum solution in the probabilistic sense only, which makes them less attractive than deterministic parallel global optimization algorithms \cite{Jones2001,Paulavicius2011,Evtushenko2013,Strongin2000}.
 As a rule, efficient deterministic algorithms are intended to solve unconstrained optimization problems.
@@
 For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^i(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$.
-The set of Peano curves thus constructed allows one to obtain $y', y''$ from $D$ for any close multidimensional images, which differ only in one coordinate, close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$.
+The set of Peano curves thus constructed allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$ for any close multidimensional images $y', y''$ from $D$, which differ only in one coordinate.
 
@@
 Previous investigations have confirmed experimentally that the index method is not inferior to well-known analogues.
-A comparison of the method with the well known DIRECT method \cite{Jones2001} in solving unconstrained optimization problems has been performed in \cite{Barkalov2015,Barkalov2016}.
+A comparison of the method with the well known DIRECT method \cite{Jones2001} in solving unconstrained optimization problems has been performed in \cite{Barkalov2015}.
 In the present study, an experimental investigation is carried out of the speedup, which is obtained by the use of the index method in combination with the two-level parallelization scheme from \cite{Sidorov2015}.
@@
 
+\bibitem{Strongin2000}
+Strongin, R.G., Sergeyev, Ya.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000)
+
 \bibitem{Balandin2011}
@@
 Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 (2017)
-
-\bibitem{Strongin2000}
-Strongin, R.G., Sergeyev, Ya.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000)
 
@@
 
-\bibitem{Barkalov2016}
-Barkalov, K., Gergel, V.: Parallel global optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016)
+%\bibitem{Barkalov2016} Barkalov, K., Gergel, V.: Parallel global optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016)
 

### 2018-05-13 2018/YSC/ysc_paper.tex
YSC: update abstract and introduction

% easychair.tex,v 3.1 2011/12/30
%
% Select appropriate paper format in your document class as
% instructed by your conference organizers. Only withtimes
% and notimes can be used in proceedings created by EasyChair
%
% The available formats are 'letterpaper' and 'a4paper' with
% the former being the default if omitted as in the example
% below.
%
\documentclass[procedia]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

% This provides the \BibTeX macro
\usepackage{doc}
\usepackage{makeidx}

\usepackage{eqnarray,amsmath}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}

\setlength{\textfloatsep}{4pt plus 1.0pt minus 1.5pt}
\setlength{\floatsep}{4pt plus 1.0pt minus 1.5pt}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% If you plan on including some algorithm specification, we recommend
% the below package. Read more details on the custom options of the
% package documentation.
%
% \usepackage{algorithm2e}

\def\procediaConference{99th Conference on Topics of
  Superb Significance (COOL 2014)}

%\makeindex

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Comparison of dimensionality reduction schemes for dervative-free global optimization
algorithms}

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair

\titlerunning{Dimension reduction schemes in direct GO}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes into the list
% defined using \institute
%
\author {
Vladislav Sovrasov
}

% Institutes for affiliations are also joined by \and,
\institute  {
  Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
  \email{sovrasov.vlad@gmail.com}\\
}
%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{Sovrasov V.}

\begin{document}

\maketitle

\keywords{Global optimization, Dimension reduction, Derivative-free algorithms, Global search algorithms}

\begin{abstract}
  A common approach to solving global optimization problems is to use
  univariate optimization algorithms in combination with dimensional reduction schemes.
  The paper considers five types of Peano-like space-filling curves (evolvents), which are used
  to reduce the dimension in the derivative-free algorithm of global optimization. The algorithm
  is univariate and developed within the framework of the information-statistical approach.
  This work is the first one, where convergence rates and implementations details
  of these five evolvents are considered together and directly compared.
\end{abstract}

%\setcounter{tocdepth}{2}
%{\small
%\tableofcontents}

%\section{To mention}
%
%Processing in EasyChair - number of pages.
%
%Examples of how EasyChair processes papers. Caveats (replacement of EC
%class, errors).

%------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
In the present paper, the algorithms for solving the multiextremal optimization problems
are considered. In the multiextremal problems, the opportunity of a reliable estimate of the global
optimum is based principally on the availability of some information on the function known
{\textit a priori} allowing relating the probable values of the optimized function to the known
values at the points of performed trials. Very often, such an information on the problem being
solved is represented in the form of suggestion that the objective function $\varphi(y)$ satisfies
Lipschitz condition with the constant $L$ not known a priori (see, for example,
\cite{Jones2009,Gablonsky2001,Evtushenko2013}). At that, the objective function could be represented by a ``black-box''-function.
Many methods destined to solving the problems of the class specified above reduce the solving
of a multidimensional problem to solving the one-dimensional subproblems implicitly (see, for
example, the methods of diagonal partitions \cite{Sergeyev2006,SergeyevKvasov2015} or
simplicial partitions \cite{Zilinskas2008,Zilinskas2014}). In the present work, we will use the
approach developed in Lobachevsky State University of Nizhni Novgorod based on the idea of
the dimensionality reduction with the use of Peano space-filling curves $y(x)$ mapping the
interval $[0,1]$ of the real axis onto an $n$-dimensional cube continuously and unambiguously.

In recent years several methods of constructing Peano-type space-filling curves (\textit{evolvents})
have been proposed \cite{strongin1978,Strongin1992,Goryachih2017,Gergel2009}. Some of them were
succesful used to build large-scale parallel algorithms \cite{stronginGergelBarkalovParGO}.
Presented paper numerically compares different evolvents in terms of convergence rates and
computational efficiency. The comparison is done by solving sets consisting of hundreds of test problems and collecting
different characteristics of convergence. As a result, practical recommendations about
the usage of each evolvent type are given. For example, some evolvents are suitable for parallel computations,
while others are quite useful in case of sequential solving of problems with time-consuming objective functions.

%------------------------------------------------------------------------------
\section{Statement of Multidimensional Global Optimization Problem}
In this paper, the core class of optimization problems, which can be solved using
univariate algorithm of global search \cite{strongin1978}, is formulated. This class involves the multidimensional global
optimization problems without constraints, which can be defined in the following way:
\begin{equation}
\label{eq:task}
\begin{array}{cr}\\
  \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
  D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
\end{array}
\end{equation}
with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function
\(\varphi(y)\) satisfies the Lipschitz condition
\begin{equation}
\label{eq:lip}
|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
\end{equation}
where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbb{R}^N\)
space.
\par
Usually, the objective function \(\varphi(y)\) is defined as a computational procedure,
according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
(let us further call such a calculation \textit{a trial}). It is supposed that this procedure
is time-consuming.
%As a result, the overall time of solving the optimization
%problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
%It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
%is highly important, since an estimate of the global minimum can be constructed on the
%basis of a finite number of computed values of the optimized function only in this case .

%------------------------------------------------------------------------------
\section{Methods of Dimension Reduction}
\subsection{Single evolvent}

Within the framework of the information-statistical global optimization theory,
the Peano space-filling curves (or evolvents) \(y(x)\) mapping the interval \([0,1]\)
onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
reduction \cite{sergeyevStronginLera2013}, \cite{strongin1978},
\cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
\par
As a result of the reduction, the initial multidimensional global optimization
problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
\begin{equation}
\label{eq:oneDimTask}
\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
\end{equation}
\par
It is important to note that this dimensionality reduction scheme transforms the % minimized
Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.
\begin{equation}
\label{eq:holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
\end{equation}
where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem
(\ref{eq:task}).
\par
The algorithms for the numerical construction of the Peano curve approximations are
given in \cite{strSergGO}.

\par
The computational scheme obtained as a result of the dimensionality reduction consists of the
following:
\begin{itemize}
  \item The optimization algorithm performs the minimization of the reduced one-dimensional
  function \(\varphi(y(x))\) from (\ref{eq:oneDimTask}),
  \item After determining the next trial point \(x\), a multidimensional image \(y\) is calculated by
using the mapping \(y(x)\),
  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the point
\(y\in D\),
  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
\end{itemize}

%------------------------------------------------------------------------------
\subsection{Shifted evolvents}
\label{sec:shifted}

One of the possible ways to overcome the negative effects of using a numerical
approximation of evolvent (it destroys the information about the neighbour points in $\mathbb{R}^N$ space,
see \cite{Strongin1992}) consists in using the multiple mappings
\begin{equation}%\label{eq:142}
Y_L(x)=\left\{y^0(x),\ y^1(x),...,\ y^L(x)\right\}
\end{equation}
instead of single Peano curve $y(x)$ (see \cite{Strongin1992,strSergGO,Strongin1991}).

Such set of evolvents can be produced by shifting the source evolvent $y^0(x)$ by $2^{-l},0
\leq l \leq L$ on each coordinate. Each evolvent has it's own corresponding hypercude $D_l=
\left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq
L$.

In Fig.~\ref{fig:shifted_ev} the image of the interval $[0,1]$ obtained by the curve $y^0(x),\
x\in [0,1],$ is shown as the dashed line. Since the hypercube $D$ from (\ref{eq:task}) is
included in the common part of the family of hypercubes $D_l$, having introduced an additional constraint
function
\begin{equation}\label{6_g0}
g_0(y)=\max\left\{\left|y_i\right| - 2^{-1}:\ 1\leq i\leq N\right\},
\end{equation}
one can present the initial hypercube $D$ in the form
\[
D=\left\{y^l(x):\; x\in [0,1],\ g_0(y^l(x))\leq 0 \right\},\ 0\leq l \leq L,
\]
i.e., $g_0(y) \leq 0$ if $y\in D$ and $g_0(y)>0$ otherwise. Consequently, any point $y \in D$
has its own preimage $x^l \in [0,1]$ for each mapping $y^l(x),\ 0\leq l\leq L$.

Thus, each evolvent $y^l(x),\ 0\leq l \leq L,$ generates its own problem of the type
(\ref{eq:task}) featured by its own extended (in comparison with $D$) search domain $D_l$
and the additional constraint with the left hand part from (\ref{6_g0})
\begin{equation}\label{6_problem_l}
\min{\left\{\varphi(y^l(x)):x\in [0,1], \; g_j(y^l(x))\leq 0, \; 0 \leq j \leq m\right\}}, \ 0 \leq l \leq L.
\end{equation}

\begin{figure}[ht]
    \centering
    \subfloat[Two shifted evolvents on the hypercubes $D_0$ and $D_1$]{{\includegraphics[width=.5\textwidth]{pictures/shifted.pdf}}\label{fig:shifted_ev}}
    %\subfloat[Hypercubes $D_l$]{{\includegraphics[width=.4\textwidth]{pictures/shifted_cube.png}}\label{fig:shifted_cube}}
    \subfloat[Two rotated evolvents on the same plane]{{\includegraphics[width=.5\textwidth]{pictures/rotated.pdf}}\label{6_fig_9}}
    \caption{Multiple evolvents built with low density }
\end{figure}

%------------------------------------------------------------------------------
\subsection{Rotated evolvents}
The application of the scheme for building the multiple evolvents (hereinafter called the shifted
evolvents or $S$-evolvents) described in Subsection \ref{sec:shifted} allows to preserve the
information on the nearness of the points in the multidimensional space and, therefore, to
provide more precise (as compared to a single evolvent) estimate of Lipschitz constant in the
search process. However, this approach has serious restrictions, which narrow the applicability
of the parallel algorithms, designed on the base of the $S$-evolvents (see the end of the section
\ref{sec:seq_comp}).

To overcome complexity of the $S$-evolvent and to preserve the information on the nearness of the points in
the $N$-dimensional space, one more scheme of building of the multiple mappings was proposed.
The building of a set of Peano curves not by the shift along the main diagonal of the hypercube
but by rotation of the evolvents around the coordinate origin is a distinctive feature of the
proposed scheme \cite{Gergel2009}.
In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case
$N=2$ are presented as an illustration.
Taking into account the initial mapping, one can conclude that current implementation of the
method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain
onto the corresponding one-dimensional intervals. Moreover, the additional constraint  $g_0(y)
\leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent. This
method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to
$2^N$) if necessary.

%\begin{figure}[t]
%  \centering
%  \includegraphics[width=0.6\linewidth]{pictures/rotated.pdf}
%  \caption{Two rotated evolvents on the same plane}
%  \label{6_fig_9}
%\end{figure}

%------------------------------------------------------------------------------
\subsection{Non-Univalent evolvent}

As it has been already mentioned above (Sec.~\ref{sec:shifted}), the loss of information on the
proximity of the points in the multidimensional space could be compensated in part by the use
of multiple mappings $Y_L(x)=\{y^1(x),...,y^L(x)\}$. However, the Peano-type curve preserves
a part of this information itself: it is not an injective mapping. Therefore, if a single image
$y(x)\in \mathbb{R}^N$ is available, on can obtain several different preimages
$t_j\in[0,1], t_j \not = x$, which could be added into the search information of the method later.

The Peano-type curve used in (\ref{eq:oneDimTask}) for the dimensionality reduction is
defined via the transition to the limit. Therefore, it cannot be computed directly. In the
numerical optimization, some approximation of this curve is used, and it is an injective piecewise-linear curve. In \cite{strongin1978} a non-univalent mapping of a uniform grid in the
interval $[0,1]$ onto a uniform grid in a hypercube $D$ has been proposed. Each
multidimensional node can have up to $2^N$ one-dimensional preimages. In
Fig.~\ref{fig:noninjective}, the grid in the $\mathbb{R}^2$ space is marked by the crosses, for
two nodes of which the corresponding one-dimensional preimages from $[0,1]$ are pointed
(marked by the squares and circles). Each node mentioned above has 3 preimages.

A potentially large number of preimages (up to $2^N$) and the inability to use the parallel
scheme for the multiple mappings are the disadvantages of the non-univalent evolvent.

%------------------------------------------------------------------------------
\subsection{Smooth evolvent}

The methods of constructing the evolvents considered in the previous paragraphs produce the
curve $y(x)$, which in not a smooth one (see Fig.~\ref{fig:shifted_ev}). The absence of
smoothness may affect the properties of the reduced one-dimensional function $\varphi(y(x))$
adversely since a smooth curve reflect the information on the growth/decay of the initial
function better. On the basis of initial algorithm of constructing the non-smooth evolvent, a
generalized algorithm allowing constructing a smooth space-filling curve has been
proposed~\cite{Goryachih2017}. As an illustration, a smooth evolvent for the two-dimensional
case is presented in Fig.~\ref{fig:smooth}.
An increased computational complexity (several times as compared to the piecewise-linear
curves) is a disadvantage of the smooth evolvent. This caused by computing of the nonlinear smooth
functions.
%At that, the number of smooth intervals and the difficulty of the
%computations of the curve increases with increasing accuracy of approximation and
%space dimension.

\begin{figure}[ht]
    \centering
    \subfloat[Smooth evolvent]{{\includegraphics[width=.6\textwidth]{pictures/smooth.pdf}}\label{fig:smooth}}
    \subfloat[Non-univalent
evolvent]{{\includegraphics[width=.4\textwidth]{pictures/noninjective.pdf}}\label{fig:noninjective}}
    \caption{Different evolvents built with low density}
\end{figure}

\section{Core Multidimensional Algorithm of Global Search}

The optimization methods applied in Globalizer to solve the reduced problem
(\ref{eq:oneDimTask}) are based on the MAGS method, which can be presented as follows ---
see \cite{strongin1978}, \cite{strSergGO}.
\par
The initial iteration of the algorithm is performed at an arbitrary point \mbox{\(x^1\in(0,1)\)}.
Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the
following rules.

\textit{Rule 1}. Renumber the points of the preceding trials by the lower indices in order of
increasing value of coordinates
$0=x_0<x_1<...<x_{k+1}=1$.

\textit{Rule 2}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq
k+1\).

\textit{Rule 4}. Determine the interval with the maximum characteristic $R(t)=\max_{1\leq i
\leq k+1}R(i)$.

\textit{Rule 5}. Execute a new trial at the point \(x^{k+1}\) located within the interval with the
maximum characteristic from the previous step
  $x^{k+1}=d(x_t)$.

The stopping condition, which terminated the trials, is defined by the inequality
$\rho_t<\varepsilon$
for the interval with the maximum characteristic from Step 4 and \(\varepsilon >0\) is the
predefined
accuracy of the optimization problem solution. If the stopping condition is not satisfied,
the index \(k\) is incremented by 1, and the new global optimization iteration is executed.

The convergence conditions and exact formulas for descision rules $R(i)$ and $d(x)$ of the
described algorithm are given, for example, in \cite{strSergGO}.

%------------------------------------------------------------------------------
\section{Results of Numerical Experiments}
%\begin{Russian}

The computational experiments have been carried out on the Lobachevsky supercomputer at
State University of Nizhni Novgorod. A computational node included 2 Intel
Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. The CPUs had 8 cores (i. e. total 16
cores were available per a node). All considered algorithms and evolvents were implemented
using C++ within the Globalizer software system~\cite{globalizerSystem}.

The comparison of the global optimization algorithms was performed by the evaluation of the
quality of solving a set of problems from some test class.
In the present paper, the test class generated by GKLS generator~\cite{Gaviano2003} was
considered. This generator allows constructing the complex multiextremal problems of various
dimensions. In the present work, the series of 100 problems from the classes of the dimensions
of 2, 3, 4, and 5 were considered.
Each class had two degrees of complexity --- \textit{Simple} and \textit{Hard}. The parameters
of the generator for the considered classes were given in Ref.~\cite{Gaviano2003}.

In order to evaluate the efficiency of an algorithm on a given set of 100 problems, we will use
the operating characteristics \cite{grishaginClass}, which are defined as a
curve, showing the dependency of number of solved problems vs the number of iterations.

%------------------------------------------------------------------------------
\subsection{Comparison of the sequential evolvents}
\label{sec:seq_comp}
%\begin{Russian}
In order to understand whether any type of evolvents listed above has an essential advantage as
compared to other ones, the operating characteristics of the index method with different types
of evolvents have been obtained for the classes GKLS 2d Simple and GKLS 3d Simple. The
global minimum was considered to be found if the algorithm generates a trial point $y^k$ in the
$\delta$-vicinity of the global minimizer, i.e. $\left\|y^k-y^\ast\right\|_\infty\leq\delta$. The size
of the vicinity was selected as $\delta = 0.01\left\|b-a\right\|_\infty$. In case of GKLS
$\delta=0.01$.

In all experiments, the evolvent construction density parameter $m=12$. The minimum value of
the reliability parameter \(r\) was found for each type of evolvents by scanning over a uniform grid
with the step \(0.1\).

On the GKLS 2d Simple class at the minimum \(r\), the non-univalent evolvent and the smooth
one provide a faster convergence (Fig.~\ref{fig:gkls2d_opt}). The same was observed at
\(r=5.0\) as well (Fig.~\ref{fig:gkls2d_acc}). In the latter case, the shifted evolvent and the
rotating one begin to lag behind the rest since the value \(r=5.0\) is too big for them.
\begin{figure}[ht]
    \centering

\subfloat[$r=5.0$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS2d_same_r_opt_pt_op.pdf}}\label{fig:gkls2d_acc}}
    \subfloat[Minimal
$r$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS2d_opt_pt_op.pdf}}\label{fig:gkls2d_opt}}
    \caption{Operating characteristics on GKLS 2d Simple class}
\end{figure}

On the GKLS 2d Simple class at the minimum \(r\), the non-univalent evolvents and multiple
ones have a considerable advantage over the single evolvent (Fig.~\ref{fig:gkls3d_opt}).
The value \(r=4.5\) is too big for the rotated evolvents and for the shifted one
(Fig.~\ref{fig:gkls3d_acc}).

\begin{figure}[ht]
    \centering

\subfloat[$r=4.5$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS3d_same_r_opt_pt_op.pdf}}\label{fig:gkls3d_acc}}
    \subfloat[Minimal
$r$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS3d_opt_pt_op.pdf}}\label{fig:gkls3d_opt}}
    \caption{Operating characteristics on GKLS 3d Simple class}
\end{figure}

\paragraph{Overhead costs when using the shifted evolvents.}
In all experiments presented above, the number of computations of the objective function from
the GKLS class was taken into account when plotting the operating characteristics. However, in
the case of the shifted evolvent, the index method solves the problem with the constraint \(g_0\)
from (\ref{6_g0}). At the points where \(g_0\) is violated, the value of the objective function is
not computed. Nevertheless, these points are stored in the search information producing the
additional computational costs. In Table~\ref{tab:shifted_g0}, the averaged numbers of calls to
\(g_0\) and to the objective function are presented. At \(L=3\), the constraint \(g_0\) was
computed almost 20 times more than the objective function \(\varphi\) i. e. \(95\%\) of the whole
search information account for the auxiliary points. Such overhead costs are acceptable when
solving the problems of small dimension with the computation costly objective functions.
However, when increasing dimensionality and total number of trials other types of evolvents are
preferred.

\begin{table}
\begin{center}
\caption{Averaged number of computations of \(g_0\) and of \(\varphi\) when solving the
problems from GKLS 3d Simple class using the shifted evolvent}
  \begin{tabular}{|l|{c}|{c}|{c}|}
    \hline
  $L$ & $calc(g_0)$ & $calc(\varphi)$ & $\frac{calc(g_0)}{calc(\varphi)}$ ratio \\
  \hline
  2 & 96247.9  & 6840.14 & 14.07\\
  \hline
  3 & 153131.0 & 7702.82 & 19.88\\
  \hline
  \end{tabular}
  \label{tab:shifted_g0}
\end{center}
\end{table}

%------------------------------------------------------------------------------
\section{Conclusions}
In the present work, 5 different Peano curve-type mappings applied to the dimensionality
reduction in the global optimization problems were considered.
From the preliminary comparison conducted in Sec.~\ref{sec:seq_comp}, one can make the
following conclusions:
\begin{itemize}
  \item the smooth evolvent and the non-univalent one demonstrate the best result in the
problems of small dimensionality and can be applied successfully in solving the problems with
the computational costly objective functions. The properties of these evolvents donï¿½t allow
developing the optimization algorithms scalable onto several cluster nodes based on these ones.
  \item the shifted evolvents introduce large overhead costs on the operation of the method due
to the requirement to adding an auxiliary functional constraint into the problem (\ref{eq:task}).
The experiments have demonstrated that up to 95\% of the search information account for the
points, in which the auxiliary constraint is computed only.
However, if the objective function is computation-costly enough, the use of these evolvents
could make sense.
  \item the rotated evolvents have provided an acceptable speed of convergence in the problems
of small dimensionality in the sequential mode. The use of these ones doesn't result in the
introduction of the auxiliary constraints that allows constructing an efficient parallel algorithm
based on these evolvents.
\end{itemize}

%------------------------------------------------------------------------------
% Index
%\printindex
%\label{sect:bib}
\bibliographystyle{unsrt}
\bibliography{ysc_paper}

%------------------------------------------------------------------------------
\end{document}

% EOF

### 2018-05-14 2018/YSC/ysc_paper.tex
YSC: fix running header
YSC: fix typos

@@
 %
-\title{Comparison of dimensionality reduction schemes for dervative-free global optimization
+\title{Comparison of dimensionality reduction schemes for derivative-free global optimization
 algorithms}
@@
 
-\titlerunning{Dimension reduction schemes in direct GO}
+\titlerunning{Dimension reduction schemes in derivative-free GO}
 
@@
 have been proposed \cite{strongin1978,Strongin1992,Goryachih2017,Gergel2009}. Some of them were
-succesful used to build large-scale parallel algorithms \cite{stronginGergelBarkalovParGO}.
+successful used to build large-scale parallel algorithms \cite{stronginGergelBarkalovParGO}.
 Presented paper numerically compares different evolvents in terms of convergence rates and
@@
 
-The convergence conditions and exact formulas for descision rules $R(i)$ and $d(x)$ of the
+The convergence conditions and exact formulas for decision rules $R(i)$ and $d(x)$ of the
 described algorithm are given, for example, in \cite{strSergGO}.
@@
 problems of small dimensionality and can be applied successfully in solving the problems with
-the computational costly objective functions. The properties of these evolvents donï¿½t allow
+the computational costly objective functions. The properties of these evolvents doesn't allow
 developing the optimization algorithms scalable onto several cluster nodes based on these ones.

### 2018-05-15 2018/ITNT/ITNT2018.tex
Extended abstract

@@
-\documentclass[a4paper]{locconf}
+\documentclass[a4paper]{jpconf}
 \usepackage{graphicx}
@@
 Nizhny Novgorod, Russia, 603950}
+
+\ead{konstantin.barkalov@itmm.unn.ru}
 
@@
 two Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM, and three NVIDIA 
-Kepler K20Õ GPUs (2688 CUDA cores, 6 Gb GDDR5). 
+Kepler K20 GPUs (2688 CUDA cores, 6 Gb GDDR5). 
 The results of experiments demonstrate that the parallel index algorithm combined with the nested optimization scheme provides a good speedup on the considered problem class.
@@
 \bibitem{Strongin2000}
-Strongin, R.G. Global optimization with non-convex constraints. Sequential 
-and parallel algorithms. / R.G. Strongin, Ya.D.  Sergeyev --- Dordrecht: 
-Kluwer Academic Publishers, 2000. --- 704 p.
+Strongin R G and Sergeyev Ya D 2000 \textit{Global Optimization with Non-Convex Constraints. Sequential 
+and Parallel Algorithms} (Dordrecht: Kluwer Academic Publishers)
 
 \bibitem{Strongin2013}
-Strongin, R.G. Parallel computing in global optimization problems / R.G. 
-Strongin, V.P. Gergel, V.A. Grishagin, K.A. Barkalov --- Moscow: Publishing 
-of the Moscow State University, 2013. --- 280 p. --- (in Russian).
+Strongin R G, Gergel V P, Grishagin V A and Barkalov K A 2013  \textit{Parallel Computing in Global Optimization Problems} (Moscow: Moscow State University Press) (in Russian).
 
 \bibitem{Globalizer}
-Sysoyev, A.V.
-Globalizer -- A parallel software system for solving global optimization problems / A.V.
-Sysoyev, K.A.
-Barkalov, V.V.
-Sovrasov, I.G.
-Lebedev, V.P.
-Gergel // Lecture Notes in Computer Science.
---- 2017.
---- Vol.
-10421.
---- P.
-492--499.
+Sysoyev A V, Barkalov K A, Sovrasov V V, Lebedev I G and Gergel V P 2017 \textit{Lecture Notes in Computer Science} \textbf{10421} 492--99
 
 \bibitem{Jones}
-Jones, D.R. The DIRECT global optimization algorithm / In: Floudas, C. A., 
-Pardalos, P. M. (eds.) The Encyclopedia of Optimization, Second Edition. --- 
-Heidelberg: Springer, 2009. --- P. 725--735.
+Floudas C A and Pardalos P M 2009 \textit{The Encyclopedia of Optimization } (Heidelberg: Springer) p 725--35
 
 \bibitem{Zilinskas}
-Paulavi\v{c}ius, R. Parallel branch and bound for global optimization with 
-combination of Lipschitz bounds / R. Paulavi\v{c}ius, J. \v{Z}ilinskas, A. 
-Grothey // Optimization Methods \& Software. --- 2011. --- Vol. 26(3). --- P. 
-487--498.
+Paulavi\v{c}ius R, \v{Z}ilinskas J and Grothey A 2011 \textit{Optim. Meth. Soft.} \textbf{26(3)} 487--98
 
 \bibitem{Evtushenko2009}
-Evtushenko, Yu.G. Parallel global optimization of functions of several 
-variables / Yu.G. Evtushenko, V.U. Malkova, A.A. Stanevichyus // 
-Computational Mathematics and Mathematical Physics. --- 2009. --- Vol. 49 
-(2). --- P. 246--260.
+Evtushenko Yu G, Malkova V U and Stanevichyus A A 2009 \textit{Comput. Math. Math. Phys.} \textbf{49(2)} 246--60.
 
 \bibitem{Evtushenko2014}
-Evtushenko, Yu.G.
-Method of non-uniform coverages to solve the multicriteria optimization problems with guaranteed accuracy / Yu.G.
-Evtushenko, M.
-A.
-Posypkin // Automation and Remote Control.
---- 2014.
---- Vol.
-75(6).
---- P.
-1025--1040.
+Evtushenko Yu G and Posypkin M A 2014 \textit{Autom. Remote Control} \textbf{75(6)} 1025--40
 
 \bibitem{Strongin2017}
-Barkalov, K.A. Test problems for parallel algorithms of constrained global optimization / K.A. Barkalov, R.G. Strongin // Lecture Notes in Computer Science. --- 2017. --- Vol. 10556. --- P. 18--33.
+Barkalov K A and Strongin R G 2017 \textit{Lecture Notes in Computer Science} \textit{10556} 18--33
 
 \bibitem{Sysoyev}
-Sysoyev, A.V. MPI implementation of dimension reduction multilevel scheme for 
-parallel solving the global optimization problems / A.V. Sysoyev, K.A. 
-Barkalov, V.P. Gergel, I.G. Lebedev // Russian Supercomputing Days: 
-Proceedings of the International Scientific Conference. --- Moscow: 
-Publishing of the Moscow State University, 2015. --- P. 61--68.
+Sysoyev A V, Barkalov K A, Gergel V P and Lebedev I G 2015 \textit{CEUR Workshop Proceedings} \textbf{1482} 61--68
 
 \bibitem{BarkalovLebedev}
-Barkalov, K.A. Solving global optimization problems on GPU / K.A. Barkalov, 
-I.G. Lebedev // Russian Supercomputing Days: Proceedings of the international 
-conference. --- Moscow: Publishing of the Moscow State University, 2016. --- 
-P. 640--650.
+Barkalov K A and Lebedev I G 2016 \textit{Communications in Computer and Information Science} \textbf{687} 224--35
 
 \bibitem{BarkalovGergel}
-Barkalov, K. Parallel global optimization on GPU / K. Barkalov K., V. Gergel 
-// Journal of Global Optimization. --- 2016. --- Vol. 66(1). --- P. 3--20.
+Barkalov K and Gergel V 2016 \textit{J. Glob. Optim.} \textbf{66(1)} 3--20
 
 \bibitem{BarkalovGergelLebedev}
-Barkalov, K. Use of Xeon Phi Coprocessor for Solving Global Optimization 
-Problems / K. Barkalov, V. Gergel, I. Lebedev //  Lecture Notes in Computer 
-Science. --- 2015. --- Vol. 9251. --- P. 307--318.
+Barkalov K, Gergel V, Lebedev I 2015 \textit{Lecture Notes in Computer Science} \textbf{9251} 307--18
 
 \bibitem{Gergel2017}
-Gergel, V. An Approach for Generating Test Problems of Constrained Global 
-Optimization / V. Gergel // Lecture Notes in Computer Science. --- 2017. --- 
-Vol. 10556. --- P. 314--319.
+Gergel V 2017 \textit{Lecture Notes in Computer Science} \textbf{10556} 314--19
 
 \bibitem{Gergel2016}
-Gergel, V. Adaptive nested optimization scheme for multidimensional global 
-search / V. Gergel, V. Grishagin, A. Gergel // Journal of Global 
-Optimization. --- 2016. --- Vol. 66(1). --- P. 35--51.
+Gergel V, Grishagin V and Gergel A 2016 \textit{J. Glob. Optim.} \textbf{66(1)} 35--51
 
 \bibitem{Sergeyev2015}
-Sergeyev, Y.D.
-(2015) A deterministic global optimization using smooth diagonal auxiliary functions / Y.D.
-Sergeyev, D.E.
-Kvasov // Communications in Nonlinear Science and Numerical Simulation.
---- 2015.
---- Vol.
-21 (1-3).
---- P.
-99--111.
+Sergeyev Y D and Kvasov D E 2015 \textit{Commun. Nonlinear Sci. Numer. Simul.} \textbf{21(1-3)} 99--111
 

### 2018-05-16 2018/YSC/ysc_paper.tex
YSC: fix grammar issue
YSC: fix typos

@@
 Such set of evolvents can be produced by shifting the source evolvent $y^0(x)$ by $2^{-l},0
-\leq l \leq L$ on each coordinate. Each evolvent has it's own corresponding hypercude $D_l=
+\leq l \leq L$ on each coordinate. Each evolvent has it's own corresponding hypercube $D_l=
 \left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq
@@
 a part of this information itself: it is not an injective mapping. Therefore, if a single image
-$y(x)\in \mathbb{R}^N$ is available, on can obtain several different preimages
+$y(x)\in \mathbb{R}^N$ is available, one can obtain several different preimages
 $t_j\in[0,1], t_j \not = x$, which could be added into the search information of the method later.
@@
 problems of small dimensionality and can be applied successfully in solving the problems with
-the computational costly objective functions. The properties of these evolvents doesn't allow
+the computational costly objective functions. The properties of these evolvents don't allow
 developing the optimization algorithms scalable onto several cluster nodes based on these ones.

### 2018-05-16 2018/RSD/MultiEvolvents/RSD.tex
RSD: add changlog for reviewers

--- 2018-04-08 2018/RSD/MultiEvolvents/RSD.tex
@@
 \documentclass[runningheads]{llncs}
-%
-%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
-\usepackage[utf8x]{inputenc}
-\usepackage[english,russian]{babel}
-\usepackage{cmap}
 %%%%%
@@
 %
-\title{Comparison of dimensionality reduction schemes for parallel global optimization algorithms
-\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150}}
+\title{Comparison of dimensionality reduction schemes for parallel global optimization
+algorithms
+\thanks{The study was supported by the Russian Science Foundation, project No 16-11-
+10150}}
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Dimension reduction schemes in parallel GO}
 % If the paper title is too long for the running head, you can set
@@
 \institute{
-Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
+Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
 \email{konstantin.barkalov@itmm.unn.ru}\\
@@
 \email{ilya.lebedev@itmm.unn.ru}\\
-\url{http://hpc-education.unn.ru/основные-напраления/глобальная-оптимизация} }
+\url{http://hpc-education.unn.ru/}
+}
 %
@@
 This work considers a parallel algorithms for solving multi-extremal optimization problems.
-Algorithms are developed within the framework of the information-statistical approach and implemented in a parallel solver “Globalizer” .
-The optimization problem is solved by reducing the multidimensional problem to a set of joint one-dimensional problems that are solved in parallel.
-Five types of Peano-type space-filling curves are employed to reduce dimension.
-The results of computational experiments carried out on several hundred test problems are discussed.
+Algorithms are developed within the framework of the information-statistical approach and
+implemented in a parallel solver Globalizer. The optimization problem is solved by reducing
+the multidimensional problem to a set of joint one-dimensional problems that are solved in
+parallel. Five types of Peano-type space-filling curves are employed to reduce dimension. The
+results of computational experiments carried out on several hundred test problems are discussed.
 
-\keywords{Global optimization \and  Dimension reduction \and Parallel algorithms  \and Multidimensional multiextremal optimization  \and Global search algorithms  \and Parallel computations }
+\keywords{Global optimization \and  Dimension reduction \and Parallel algorithms  \and
+Multidimensional multiextremal optimization  \and Global search algorithms  \and Parallel
+computations }
 \end{abstract}
@@
 \section{Introduction}\label{sec:intro}
+%\begin{Russian}
 
+In the present paper, the parallel algorithms for solving the multiextremal optimization problems
+are considered. In the multiextremal problems, the opportunity of reliable estimate of the global
+optimum is based principally on the availability of some information on the function known
+{\textit a priori} allowing relating the probable values of the optimized function to the known
+values at the points of performed trials. Very often, such an information on the problem being
+solved is represented in the form of suggestion that the objective function $\varphi(y)$ satisfies
+Lipschitz condition with the constant $L$ not known a priori (see, for example,
+\cite{Jones,Gablonsky,Evtushenko}). At that, the objective function could be defined by a
+program code i. e. could represent a ``black-box''-function. Such problems are presented in the
+applications widely (problems of optimal design of objects and technological processes in
+various fields of technology, problems of model fitting according to observed data in scientific
+research, etc.).
+
+Many methods destined to solving the problems of the class specified above reduce the solving
+of a multidimensional problem to solving the one-dimensional subproblems implicitly (see, for
+example, the methods of diagonal partitions \cite{Sergeyev2006,SergeyevKvasov2015} or
+simplicial partitions \cite{Zilinskas2008,Zilinskas2014}). In the present work, we will use the
+approach developed in Lobachevsky State University of Nizhni Novgorod based on the idea of
+the dimensionality reduction with the use of Peano space-filling curves $y(x)$ mapping the
+interval $[0,1]$ of the real axis onto an $n$-dimensional cube continuously and unambiguously.
+
+Several methods of constructing the evolvents approximating the theoretical Peano curve have
+been proposed in \cite{strongin1978,Strongin1992,Goryachih2017,Gergel2009}. These
+methods were implemented in the Globalizer software system \cite{globalizerSystem}. The goal
+of the present study was comparing the properties of the evolvents and the selecting the most
+suitable ones for the use in the parallel global optimization algorithms.
+
+%\end{Russian}
 %------------------------------------------------------------------------------
@@
 \end{equation}
-with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function \(\varphi(y)\) satisfies the Lipschitz condition
+with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function
+\(\varphi(y)\) satisfies the Lipschitz condition
 \begin{equation}
@@
 \end{equation}
-where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbb{R}^N\) space.
+where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbb{R}^N\)
+space.
 \par
-Usually, the minimized function \(\varphi(y)\) is defined as a computational procedure,
+Usually, the objective function \(\varphi(y)\) is defined as a computational procedure,
 according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
 (let us further call such a calculation \textit{a trial}). It is supposed that this procedure
-is a time-consuming one. As a result, the overall time of solving the optimization
-problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
-It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
-is highly important, since an estimate of the global minimum can be constructed on the
-basis of a finite number of computed values of the optimized function only in this case .
+is time-consuming.
+%As a result, the overall time of solving the optimization
+%problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
+%It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
+%is highly important, since an estimate of the global minimum can be constructed on the
+%basis of a finite number of computed values of the optimized function only in this case .
 
@@
 onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
-reduction \cite{sergeyevStronginLera2013}, \cite{strongin1978}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
+reduction \cite{sergeyevStronginLera2013}, \cite{strongin1978},
+\cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
 \par
@@
 \par
-It is important to note that this dimensionality reduction scheme transforms the minimized
+It is important to note that this dimensionality reduction scheme transforms the % minimized
 Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
@@
 where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
-constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem (\ref{eq:task}).
+constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem
+(\ref{eq:task}).
 \par
@@
 \par
-The computational scheme obtained as a result of the dimensionality reduction consists of the following:
+The computational scheme obtained as a result of the dimensionality reduction consists of the
+following:
 \begin{itemize}
@@
   function \(\varphi(y(x))\) from (\ref{eq:oneDimTask}),
-  \item After determining the next trial point \(x\), a multidimensional image \(y\) is calculated by using the
-mapping \(y(x)\),
-  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the point \(y\in D\),
+  \item After determining the next trial point \(x\), a multidimensional image \(y\) is calculated by
+using the mapping \(y(x)\),
+  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the point
+\(y\in D\),
   \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
@@
 
-The reduction of the multidimensional problems to the one-dimensional ones using the Peano curves has such important properties as the continuity and the uniform boundedness of the function differences for limited variation of argument.
-However, a partial loss of information on the nearness of the points in the multidimensional space takes place since a point $x \in [0,1]$ has the left and the right neighbors only while the corresponding point $y(x) \in D \subset R^N$ has the neighbors in $2N$ directions.
-As a result, when using the mappings like Peano curve the images $y',\ y'',$ which are close to each other in the $N$-dimensional space can correspond to the preimages $x',\ x'',$ which can be far away from each other in the interval $[0,1]$.
-This property results in the excess computations since several limit points $x',\ x''$ of the trial sequence generated by the index method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional space.
-
-One of the possible ways to overcome this disadvantage consists in using the multiple mappings
+One of the possible ways to overcome the negative effects of using a numerical
+approximation of evolvent (it destroys the information about the neighbour points in $\mathbb{R}^N$ space,
+see \cite{Strongin1992}) consists in using the multiple mappings
 \begin{equation}%\label{eq:142}
@@
 \end{equation}
-instead of single Peano curve $y(x)$ (see \cite{Strongin1991,Strongin1992,strSergGO}).
+instead of single Peano curve $y(x)$ (see \cite{Strongin1992,strSergGO,Strongin1991}).
 
-Such set of evolvents can be produced by shifting the source evolvent $y^0(x)$ by $2^{-l},0 \leq l \leq L$ on each coordinate.
-Each evolvent has it's own corresponding hypercude $D_l= \left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq L$.
+Such set of evolvents can be produced by shifting the source evolvent $y^0(x)$ by $2^{-l},0
+\leq l \leq L$ on each coordinate. Each evolvent has it's own corresponding hypercube $D_l=
+\left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq
+L$.
 
-In Fig.~\ref{fig:shifted_ev} the image of the interval $[0,1]$ obtained by the curve $y^0(x),\ x\in [0,1],$ is shown as the dashed line.
-Since the hypercube $D$ from (\ref{eq:task}) is included in the common part of the family of hypercubes $D_l$ (the boundaries of hypercube $D$ are highlighted in Fig.~\ref{fig:shifted_cube}), having introduced an additional constraint function
+In Fig.~\ref{fig:shifted_ev} the image of the interval $[0,1]$ obtained by the curve $y^0(x),\
+x\in [0,1],$ is shown as the dashed line. Since the hypercube $D$ from (\ref{eq:task}) is
+included in the common part of the family of hypercubes $D_l$, having introduced an additional constraint
+function
 \begin{equation}\label{6_g0}
@@
 \]
-i.e., $g_0(y) \leq 0$ if $y\in D$ and $g_0(y)>0$ otherwise. Consequently, any point $y \in D$ has its own preimage $x^l \in [0,1]$ for each mapping $y^l(x),\ 0\leq l\leq L$.
+i.e., $g_0(y) \leq 0$ if $y\in D$ and $g_0(y)>0$ otherwise. Consequently, any point $y \in D$
+has its own preimage $x^l \in [0,1]$ for each mapping $y^l(x),\ 0\leq l\leq L$.
 
-Thus, each evolvent $y^l(x),\ 0\leq l \leq L,$ generates its own problem of the type (\ref{eq:task}) featured by its own extended (in comparison with $D$) search domain $D_l$ and the additional constraint with the left hand part from (\ref{6_g0})
+Thus, each evolvent $y^l(x),\ 0\leq l \leq L,$ generates its own problem of the type
+(\ref{eq:task}) featured by its own extended (in comparison with $D$) search domain $D_l$
+and the additional constraint with the left hand part from (\ref{6_g0})
 \begin{equation}\label{6_problem_l}
@@
     \centering
-    \subfloat[Two mappings]{{\includegraphics[width=.6\textwidth]{pictures/shifted.pdf}}\label{fig:shifted_ev}}
-    \subfloat[Hypercubes $D_l$]{{\includegraphics[width=.4\textwidth]{pictures/shifted_cube.png}}\label{fig:shifted_cube}}
-    \caption{Multiple mappings}
+    \subfloat[Two shifted evolvents on the hypercubes $D_0$ and $D_1$]{{\includegraphics[width=.5\textwidth]{pictures/shifted.pdf}}\label{fig:shifted_ev}}
+    %\subfloat[Hypercubes $D_l$]{{\includegraphics[width=.4\textwidth]{pictures/shifted_cube.png}}\label{fig:shifted_cube}}
+    \subfloat[Two rotated evolvents on the same plane]{{\includegraphics[width=.5\textwidth]{pictures/rotated.pdf}}\label{6_fig_9}}
+    \caption{Multiple evolvents built with low density }
 \end{figure}
@@
 \subsection{Rotated evolvents}
-The application of the scheme for building the multiple evolvents (hereinafter called the shifted evolvents or $S$-evolvents) described in Subsection \ref{sec:shifted} allows to preserve the information on the nearness of the points in the multidimensional space and, therefore, to provide more precise (as compared to a single evolvent) estimate of Lipschitz constant in the search process.
-However, this approach has serious restrictions, which narrow the applicability of the parallel algorithms, designed on the base of the $S$-evolvents (see te end of the section \ref{sec:seq_comp}).
+The application of the scheme for building the multiple evolvents (hereinafter called the shifted
+evolvents or $S$-evolvents) described in Subsection \ref{sec:shifted} allows to preserve the
+information on the nearness of the points in the multidimensional space and, therefore, to
+provide more precise (as compared to a single evolvent) estimate of Lipschitz constant in the
+search process. However, this approach has serious restrictions, which narrow the applicability
+of the parallel algorithms, designed on the base of the $S$-evolvents (see the end of the section
+\ref{sec:seq_comp}).
 
-To overcome the $S$-evolvent and to preserve the information on the nearness of the points in the $N$-dimensional space, a novel scheme of building of the multiple mappings is proposed.
-The building of a set of Peano curves not by the shift along the main diagonal of the hypercube but by rotation of the evolvents around the coordinate origin is a distinctive feature of the proposed scheme \cite{Gergel2009}.
-In the initial non-rotated mapping for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other.
-In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer.
-The evolvents generated according to this scheme are called the rotated evolvents or $R$-evolvents.
+To overcome complexity of the $S$-evolvent and to preserve the information on the nearness of the points in
+the $N$-dimensional space, one more scheme of building of the multiple mappings was proposed.
+The building of a set of Peano curves not by the shift along the main diagonal of the hypercube
+but by rotation of the evolvents around the coordinate origin is a distinctive feature of the
+proposed scheme \cite{Gergel2009}.
+In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case
+$N=2$ are presented as an illustration.
+Taking into account the initial mapping, one can conclude that current implementation of the
+method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain
+onto the corresponding one-dimensional intervals. Moreover, the additional constraint  $g_0(y)
+\leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent. This
+method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to
+$2^N$) if necessary.
 
-In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case $N=2$ are presented as an illustration.
-Taking into account the initial mapping, one can conclude that current implementation of the method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals.
-Moreover, the additional constraint  $g_0(y) \leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent.
-This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.6\linewidth]{pictures/rotated_1.pdf}
-  \caption{Two rotated evolvents on the same plane}
-  \label{6_fig_9}
-\end{figure}
+%\begin{figure}[t]
+%  \centering
+%  \includegraphics[width=0.6\linewidth]{pictures/rotated.pdf}
+%  \caption{Two rotated evolvents on the same plane}
+%  \label{6_fig_9}
+%\end{figure}
 
 %------------------------------------------------------------------------------
-\subsection{Non injective evolvent}
+\subsection{Non-Univalent evolvent}
+%\begin{Russian}
 
-Как уже было сказано в секции \ref{sec:shifted}, потеря информации о близости точек в многомерном пространстве может быть частично скомпенсирована использованием множественных отображений $Y_L(x)=\{y^1(x),...,y^L(x)\}$.
-Однако, сама по себе кривая типа Пеано сохраняет в себе часть этой информации: она не является инъективым отображением, поэтому имея один образ $y(x)\in \mathbb{R}^N$, можно получить несколько несколько отличных $x$ прообразов $t_j\in[0,1], t_j \not = x$, которые затем могут быть добавлены в поисковую информацию индексного метода.
+As it has been already mentioned above (Sec.~\ref{sec:shifted}), the loss of information on the
+proximity of the points in the multidimensional space could be compensated in part by the use
+of multiple mappings $Y_L(x)=\{y^1(x),...,y^L(x)\}$. However, the Peano-type curve preserves
+a part of this information itself: it is not an injective mapping. Therefore, if a single image
+$y(x)\in \mathbb{R}^N$ is available, one can obtain several different preimages
+$t_j\in[0,1], t_j \not = x$, which could be added into the search information of the method later.
 
-Кривая типа Пеано, используемая в (\ref{eq:oneDimTask}) для редукции размерности, определяется через предельный переход,
-поэтому не может быть вычислена непосредтвенно. При численной оптимизации используется некоторое её приближение, являющееся
-инъективной кусочно-линейной кривой.
-В \cite{strongin1978} было предложено неинъективное отображение равномерной сетки на отрезка $[0,1]$ на равномерную сетку в гиперкубе $D$.
-Каждый многомерный узел может иметь до $2^N$ одномерных прообразов.
-На рис.
-\ref{fig:noninjective} изображена кривая с самопересечениями, полученная при соединении узлов грубой многомерной сетки в порядке следования их прообразов из отрезка $[0,1]$, а также отмечана точка, имеющая 3 прообраза.
+The Peano-type curve used in (\ref{eq:oneDimTask}) for the dimensionality reduction is
+defined via the transition to the limit. Therefore, it cannot be computed directly. In the
+numerical optimization, some approximation of this curve is used, and it is an injective piecewise-linear curve. In \cite{strongin1978} a non-univalent mapping of a uniform grid in the
+interval $[0,1]$ onto a uniform grid in a hypercube $D$ has been proposed. Each
+multidimensional node can have up to $2^N$ one-dimensional preimages. In
+Fig.~\ref{fig:noninjective}, the grid in the $\mathbb{R}^2$ space is marked by the crosses, for
+two nodes of which the corresponding one-dimensional preimages from $[0,1]$ are pointed
+(marked by the squares and circles). Each node mentioned above has 3 preimages.
 
-Недостатком неинъективной развёртки является потенциально большое количетво прообразов (до $2^N$) и невозможность использования параллельной схемы для множественных отображений из секции \ref{sec:parallel_evolvents}.
+A potentially large number of preimages (up to $2^N$) and the inability to use the parallel
+scheme for the multiple mappings form Sec.~\ref{sec:parallel_evolvents} are the disadvantages
+of the non-univalent evolvent.
 
@@
 
+The methods of constructing the evolvents considered in the previous paragraphs produce the
+curve $y(x)$, which in not a smooth one (see Fig.~\ref{fig:shifted_ev}). The absence of
+smoothness may affect the properties of the reduced one-dimensional function $\varphi(y(x))$
+adversely since a smooth curve reflect the information on the growth/decay of the initial
+function better. On the basis of initial algorithm of constructing the non-smooth evolvent, a
+generalized algorithm allowing constructing a smooth space-filling curve has been
+proposed~\cite{Goryachih2017}. As an illustration, a smooth evolvent for the two-dimensional
+case is presented in Fig.~\ref{fig:smooth}.
+An increased computational complexity (several times as compared to the piecewise-linear
+curves) is a disadvantage of the smooth evolvent. This caused by computing of the nonlinear smooth
+functions.
+%At that, the number of smooth intervals and the difficulty of the
+%computations of the curve increases with increasing accuracy of approximation and
+%space dimension.
+
+%\end{Russian}
+
 \begin{figure}[ht]
@@
     \subfloat[Smooth evolvent]{{\includegraphics[width=.6\textwidth]{pictures/smooth.pdf}}\label{fig:smooth}}
-    \subfloat[Numerical non injective evolvent]{{\includegraphics[width=.4\textwidth]{pictures/noninjective.pdf}}\label{fig:noninjective}}
-    \caption{Different evolvents}
+    \subfloat[Non-univalent
+evolvent]{{\includegraphics[width=.4\textwidth]{pictures/noninjective.pdf}}\label{fig:noninjective}}
+    \caption{Different evolvents built with low density}
 \end{figure}
@@
 \section{Parallel Computations for Solving Global Optimization Problems.}
-\subsection{Core Multidimensional Algorithm of Global Search}
+\subsection{Core Multidimensional Algorithm of Global Search (MAGS)}
 
-The optimization methods applied in Globalizer to solve the reduced problem (\ref{eq:oneDimTask}) are based on the MAGS method, which can be presented as follows --- see \cite{strongin1978}, \cite{strSergGO}.
+The optimization methods applied in Globalizer to solve the reduced problem
+(\ref{eq:oneDimTask}) are based on the MAGS method, which can be presented as follows ---
+see \cite{strongin1978}, \cite{strSergGO}.
 \par
@@
 Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
-The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the following rules.
+The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the
+following rules.
 
-\textit{Rule 1}. Renumber the points of the preceding trials by the lower indices in order of increasing value of coordinates
+\textit{Rule 1}. Renumber the points of the preceding trials by the lower indices in order of
+increasing value of coordinates
 $0=x_0<x_1<...<x_{k+1}=1$.
 
-\textit{Rule 2}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq k+1\).
+\textit{Rule 2}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq
+k+1\).
 
-\textit{Rule 4}. Determine the interval with the maximum characteristic $R(t)=\max_{1\leq i \leq k+1}R(i)$.
+\textit{Rule 4}. Determine the interval with the maximum characteristic $R(t)=\max_{1\leq i
+\leq k+1}R(i)$.
 
-\textit{Rule 5}. Execute a new trial at the point \(x^{k+1}\) located within the interval with the maximum characteristic from the previous step
+\textit{Rule 5}. Execute a new trial at the point \(x^{k+1}\) located within the interval with the
+maximum characteristic from the previous step
   $x^{k+1}=d(x_t)$.
 
-The stopping condition, which terminated the trials, is defined by the inequality $\rho_t<\varepsilon$
-for the interval with the maximum characteristic from Step 4 and \(\varepsilon >0\) is the predefined
+The stopping condition, which terminated the trials, is defined by the inequality
+$\rho_t<\varepsilon$
+for the interval with the maximum characteristic from Step 4 and \(\varepsilon >0\) is the
+predefined
 accuracy of the optimization problem solution. If the stopping condition is not satisfied,
@@
 
-The convergence conditions and exact formulas for descision rules $R(i)$ and $d(x)$ of the described algorithm are given, for example, in \cite{strSergGO}.
+The convergence conditions and exact formulas for descision rules $R(i)$ and $d(x)$ of the
+described algorithm are given, for example, in \cite{strSergGO}.
 
@@
 \label{sec:parallel_evolvents}
-Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the problems
+Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the
+problems
 \[
@@
 \]
-on a set of intervals $[0,1]$ by the index method.
-Each one-dimensional problem is solved on a separate processor.
-The trial results at the point \(x^k\) obtained for the problem being solved by particular processor are interpreted as the results of the trials in the rest problems (in the corresponding points \(x^(k_1 ),\dots,x^(k_S))\).
-In this approach, a trial at the point \(x^k \in [0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of operations.
+on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a
+separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
+particular processor are interpreted as the results of the trials in the rest problems (in the
+corresponding points \(x^{k_1},\dots,x^{k_S})\). In this approach, a trial at the point \(x^k \in
+[0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
+operations.
 \par
@@
 \par
-2. Inform the rest of processors about the start of the trial execution at the point\( y^k\) (the blocking of the point \(y^k\) ).
+2. Inform the rest of processors about the start of the trial execution at the point \( y^k\) (the
+blocking of the point \(y^k\) ).
 \par
-3.
-Determine the preimages \(x{}^{k_s}  \in [0,1], 1≤s≤S\), of the point \(y^k\) and interpret the trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points \(x{}^{k_1} ,…,x{}^{k_s} \)
+3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
+trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
+\(x{}^{k_1} ,\dots,x{}^{k_s} \)
 \par
@@
 
-The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the sequential algorithm (except the method of the trial execution).
-Each processor has its own copy of the software realizing the computations of the problem functions and the decision rule of the index algorithm.
-For the organization of the interactions among the processors, the queues are created on each processor, where the processors store the information on the executed iterations in the form of the tuples: the processor number \(s\), the trial point \(x{}^{k_s}\).
+The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the
+sequential algorithm (except the method of the trial execution). Each processor has its own copy
+of the software realizing the computations of the problem functions and the decision rule of the
+index algorithm. For the organization of the interactions among the processors, the queues are
+created on each processor, where the processors store the information on the executed iterations
+in the form of the tuples: the processor number \(s\), the trial point \(x{}^{k_s}\).
 \par
-The proposed parallelization scheme was implemented with the use of MPI technology.
-Main features of implementation consist in the following.
-A separate MPI-process is created for each of \(S\) one-dimensional problems being solved, usually, one process per one processor employed.
-Each process can use p threads, usually one thread per an accessible core.
+The proposed parallelization scheme was implemented with the use of MPI technology. Main
+features of implementation consist in the following. A separate MPI-process is created for each
+of \(S\) one-dimensional problems being solved, usually, one process per one processor
+employed. Each process can use $p$ threads, usually one thread per an accessible core.
 \par
-At every iteration of the method, the process with the index \(s,0\leqslant s< S\) performs p trials in parallel at the points \(x^{(s+i_S)},0\leqslant i<p\).
-At that, each process stores all \(S_p\) points, and an attribute indicating whether this point is blocked by another process or not is stored for each point.
-Let us remind that the point is blocked if the process starts the execution of a trial at this point.
+At every iteration of the method, the process with the index \(s,0\leqslant s< S\) performs p
+trials in parallel at the points \(x^{(s+iS)},0\leqslant i<p\). At that, each process stores all
+\(S_p\) points, and an attribute indicating whether this point is blocked by another process or
+not is stored for each point. Let us remind that the point is blocked if the process starts the
+execution of a trial at this point.
 \par
-At every iteration of the algorithm, operating within the \(s\)-th process, determines the coordinates of p «its own» trial points.
-Then, the interchange of the coordinates of images of the trial points \(y^{(s+i_S)},0\leqslant i<p, 0\leqslant s< S\) is performed (from each process to each one).
-After that, the preimages \(x^{(q+i_S)},0\leqslant q<S,q\not=s\) of the points received by the \(s\)-th process from the neighbor ones are determined with the use of the evolvent \(y^s (x)\).
-The points blocked within the \(s\)-th process will correspond to the preimages obtained.
-Then, each process performs the trials at the non-blocked points, the computations are performed in parallel using OpenMP.
-The results of the executed trials (the index of the point, the computed values of the problem functions, and the attribute of unblocking of this point) are transferred to all rest processes.
-All the points are added to the search information database, and the transition to the next iteration is performed.
-\par
+At every iteration of the algorithm, operating within the \(s\)-th process, determines the
+coordinates of \(p\) ``its own" trial points. Then, the interchange of the coordinates of images of
+the trial points \(y^{(s+iS)},0\leqslant i<p, 0\leqslant s< S\) is performed (from each process to
+each one). After that, the preimages \(x^{(q+iS)},0\leqslant q<S,q\not=s\) of the points
+received by the \(s\)-th process from the neighbor ones are determined with the use of the
+evolvent \(y^s (x)\). The points blocked within the \(s\)-th process will correspond to the
+preimages obtained. Then, each process performs the trials at the non-blocked points, the
+computations are performed in parallel using OpenMP. The results of the executed trials (the
+index of the point, the computed values of the problem functions, and the attribute of
+unblocking of this point) are transferred to all rest processes. All the points are added to the
+search information database, and the transition to the next iteration is performed.
 
@@
 \section{Results of Numerical Experiments}
+%\begin{Russian}
+
 The computational experiments have been carried out on the Lobachevsky supercomputer at
-State University of Nizhny Novgorod. A computational node included 2 Intel
-Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. The CPUs had 8 cores (i. e. total 16 cores
-were available per a node). Все рассматриваемые алгортимы и развёртки были реализованы на языке C++ в рамках программной системы Globalizer\cite{globalizerSystem}.
-Для реализации параллелизма на одном узле использована технология OpenMP, а для параллелизма на несколько узлов --- стандарт MPI.
+State University of Nizhni Novgorod. A computational node included 2 Intel
+Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. The CPUs had 8 cores (i. e. total 16
+cores were available per a node). All considered algorithms and evolvents were implemented
+using C++ within the Globalizer software system~\cite{globalizerSystem}.
+In order to enable the parallelism,
+OpenMP was used on a single node, and MPI was used for the parallelization on
+several nodes.
 
-Сравнение алгоритмов глобальной оптимизации проведено путём оценки качества решения алгоритмами выборки задач из некоторого тестового класса.
-В данной статье рассматривается тестовый класс, порождаемый генератором GKLS \cite{Gaviano2003}.
-Данные генератор позволяет конструировать сложные многоэкстремальные задачи различной размерности.
-В работе рассматриваются выборки по 100 задач классов размерности 2, 3, 4, 5.
-Каждый класс имеет две степени сложности --- \textit{Simple} и \textit{Hard}.
-Параметры генератора для рассматриваемых классов приведены в \cite{Gaviano2003}.
+The comparison of the global optimization algorithms was performed by the evaluation of the
+quality of solving a set of problems from some test class.
+In the present paper, the test class generated by GKLS (Gaviano, Kvasov, Lera, Sergeyev) generator~\cite{Gaviano2003} was
+considered. The generator creates objective functions by distorting a convex quadratic function by polynomials in order to introduce local minima.
+Thus, GKLS allows constructing the complex multiextremal problems of various dimensions.
+In the present work, the series of 100 problems from the classes of the dimensions
+of 2, 3, 4, and 5 were considered.
+Each class had two degrees of complexity --- \textit{Simple} and \textit{Hard}. These the classes have different
+radius of the attraction region of the global minimizer (\textit{Hard} has smaller region) and distance from the global minimizer to the vertex of the quadratic function (\textit{Simple} has smaller distance).
+The parameters of the generator for the considered classes were given in Ref.~\cite{Gaviano2003}.
 
-In order to evaluate the efficiency of an algorithm on a given set of 100 problems, we will use the
-operating characteristics \cite{grishaginClass}, which are defined by a set of
-points on the \((K, P)\) plane where \(K\) is the average number of search trials
-conducted before satisfying the termination condition when minimizing a function
-from a given class, and \(P\) is the proportion of problems solved successfully.
-If at a given \(K\), the operating characteristic of a method goes higher than one
-from another method, it means that at fixed search costs, the former method has a
-greater probability of finding the solution. If some value of \(P\) is fixed, and the
-characteristic of a method goes to the left from that of another method, the former
-method requires fewer resources to achieve the same reliability.
+%\end{Russian}
+
+In order to evaluate the efficiency of an algorithm on a given set of 100 problems, we will use
+the operating characteristics \cite{grishaginClass}, which are defined as a
+curve, showing the dependency of number of solved problems vs the number of iterations.
 
 %------------------------------------------------------------------------------
-\subsection{Сравнение последовательных разверток}
+\subsection{Comparison of the sequential evolvents}
 \label{sec:seq_comp}
-С целью понять, обладает ли какой-либо из перечисленных ранее типов развёрток существенным преимуществом над другими, были построены операционные характеристики индексного метода с различными типами развёрток на классах GKLS 2d Simple и GKLS 3d Simple.
-The global minimum was considered to be found if the algorithm generates a trial point $y^k$ in the $\delta$-vicinity of the global minimizer, i.e.
-$\left\|y^k-y^\ast\right\|_\infty\leq\delta$.
-The size of the vicinity was selected as $\delta = 0.01\left\|b-a\right\|_\infty$.
-In case of GKLS $\delta=0.01$.
+%\begin{Russian}
+In order to understand whether any type of evolvents listed above has an essential advantage as
+compared to other ones, the operating characteristics of the index method with different types
+of evolvents have been obtained for the classes GKLS 2d Simple and GKLS 3d Simple. The
+global minimum was considered to be found if the algorithm generates a trial point $y^k$ in the
+$\delta$-vicinity of the global minimizer, i.e. $\left\|y^k-y^\ast\right\|_\infty\leq\delta$. The size
+of the vicinity was selected as $\delta = 0.01\left\|b-a\right\|_\infty$. In case of GKLS
+$\delta=0.01$.
 
-Во всех экспериментах параметр плотности построения развёрток $m=12$.
-Минимальное значение пераметра надёжности \(r\) было найдено для каждого типа развёртки перебором по равномерной сетке с шагом \(0.1\).
+In all experiments, the evolvent construction density parameter $m=12$. The minimum value of
+the reliability parameter \(r\) was found for each type of evolvents by scanning over a uniform grid
+with the step \(0.1\).
 
-На классе GKLS 2d Simple при минимальном \(r\) неинъективная и гладкая развёртка обеспечивают более быструю сходимость (рис.
-\ref{fig:gkls2d_opt}).
-То же самое, наблюдается и при \(r=5.0\) (рис.
-\ref{fig:gkls2d_acc}).
-В последнем случае сдвиговая и вращаемая развёртки начинают отставать от остальных, т.к.
-значение \(r=5.0\) является завышенным для них.
+On the GKLS 2d Simple class at the minimum \(r\), the non-univalent evolvent and the smooth
+one provide a faster convergence (Fig.~\ref{fig:gkls2d_opt}). The same was observed at
+\(r=5.0\) as well (Fig.~\ref{fig:gkls2d_acc}). In the latter case, the shifted evolvent and the
+rotating one begin to lag behind the rest since the value \(r=5.0\) is too big for them.
 \begin{figure}[ht]
     \centering
+
 \subfloat[$r=5.0$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS2d_same_r_opt_pt_op.pdf}}\label{fig:gkls2d_acc}}
-    \subfloat[Minimal $r$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS2d_opt_pt_op.pdf}}\label{fig:gkls2d_opt}}
+    \subfloat[Minimal
+$r$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS2d_opt_pt_op.pdf}}\label{fig:gkls2d_opt}}
     \caption{Operating characteristics on GKLS 2d Simple class}
@@
 
-На классе GKLS 2d Simple при минимальном \(r\) неинъективная и множественные развёртки имеют значительное преимущество над единственной развёрткой (рис.
-\ref{fig:gkls3d_opt}).
-Значение \(r=4.5\) является завышенным для вращаемой и сдвиговой развёрток (рис.
-\ref{fig:gkls3d_acc}).
+On the GKLS 2d Simple class at the minimum \(r\), the non-univalent evolvents and multiple
+ones have a considerable advantage over the single evolvent (Fig.~\ref{fig:gkls3d_opt}).
+The value \(r=4.5\) is too big for the rotated evolvents and for the shifted one
+(Fig.~\ref{fig:gkls3d_acc}).
 
@@
     \centering
+
 \subfloat[$r=4.5$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS3d_same_r_opt_pt_op.pdf}}\label{fig:gkls3d_acc}}
-    \subfloat[Minimal $r$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS3d_opt_pt_op.pdf}}\label{fig:gkls3d_opt}}
+    \subfloat[Minimal
+$r$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS3d_opt_pt_op.pdf}}\label{fig:gkls3d_opt}}
     \caption{Operating characteristics on GKLS 3d Simple class}
@@
 
-\paragraph{Накладные расходы при использовании сдвиговой развёртки.}
-Во всех представленных выше экспериментах при построении операционной характеристики учитывалось количество вычислений целевой функции из класса GKLS, однако в случае сдвиговой развёртки индексный метод решает задачу с ограничением \(g_0\) из (\ref{6_g0}).
-В точках, где \(g_0\) нарушено, значение целевой функции не вычисляется.
-Эти точки, тем не менее, хранятся в поисковой информации, создавая дополнительные расходы вычислительных ресурсов.
-В таблице \ref{tab:shifted_g0} приведено среднее количество обращение к \(g_0\) и целевой функции.
-При \(L=3\) ограниение \(g_0\) вычисляется почти в 20 раз чаще, чем целевая функция \(\varphi\), т.
-е.
-\(95\%\) всей поисковой информации приходится на вспомогательные точки.
-Такие накладные расходы приемлемы при решении задач малой размерности с трудоёмкими целевыми функциями, но при росте размерности и общего количества испытаний выгоднее использовать другие типы развёрток.
+\paragraph{Overhead costs when using the shifted evolvents.}
+In all experiments presented above, the number of computations of the objective function from
+the GKLS class was taken into account when plotting the operating characteristics. However, in
+the case of the shifted evolvent, the index method solves the problem with the constraint \(g_0\)
+from (\ref{6_g0}). At the points where \(g_0\) is violated, the value of the objective function is
+not computed. Nevertheless, these points are stored in the search information producing the
+additional computational costs. In Table~\ref{tab:shifted_g0}, the averaged numbers of calls to
+\(g_0\) and to the objective function are presented. At \(L=3\), the constraint \(g_0\) was
+computed almost 20 times more than the objective function \(\varphi\) i. e. \(95\%\) of the whole
+search information account for the auxiliary points. Such overhead costs are acceptable when
+solving the problems of small dimension with the computation costly objective functions.
+However, when increasing dimensionality and total number of trials other types of evolvents are
+preferred.
 
@@
 \begin{center}
-\caption{Среднее количество вычислений \(g_0\) и \(\varphi\) при решении задач класса GKLS 3d Simple с помощью сдвиговой развёртки}
+\caption{Averaged number of computations of \(g_0\) and of \(\varphi\) when solving the
+problems from GKLS 3d Simple class using the shifted evolvent}
   \begin{tabular}{|l|{c}|{c}|{c}|}
@@
 %------------------------------------------------------------------------------
-\subsection{Параллельные вращаемые развертки}
+\subsection{Parallel rotated evolvents}
 \label{sec:results_parallel}
-Чтобы оценить эффективность параллельного алгортима из секции \ref{sec:parallel_evolvents} были проведены вычислительные эксперименты на классах GKLS 4d (Hard, Simple) и GKLS 5d (Hard, Simple).
-Значения \(r\) во всех экспериментах равно 5.0, размер \(\delta\)-окрестности известного решения увеличен до 0.3.
-При решении серий задач использовалось до 8 узлов кластера и до 32 вычислительных потоков на каждом узле.
+In order to evaluate the efficiency of the parallel algorithm from
+Sec.~\ref{sec:parallel_evolvents}, the numerical experiments on the GKLS 4d (Hard, Simple)
+classes and on the GKLS 5d (Hard, Simple) ones were conducted. The value of \(r\) in all
+experiments was equal to 5.0, the size of the \(\delta\)-vicinity of the known solution was
+increased up to 0.3. When solving the series of problems, up to 8 cluster nodes and up to 32
+computational threads on each node were employed.
 
-В таблице \ref{tab:iterations} приведено среднее количество итераций при решении 100 задач из каждого из рассматриваемых классов.
-При увеличении числа узлов и числа потоков на каждом узле количество итераций заметно сокращается (за исключением класса GKLS 4d Simple при переходе с 1 узла на 4 узла в однопоточном режиме).
+In Table~\ref{tab:iterations}, an averaged number of iterations when solving 100 problems from
+each considered class is presented.
+The number of iterations is reduced considerably with increasing the number of nodes and the
+number of threads on each node (except the GKLS 4d Simple class at the transition from 1
+node to 4 ones in the single thread mode).
 
@@
   \centering
-  \caption{Averaged numbers of iterations executed by the parallel algorithm for solving the test optimization problems}
+  \caption{Averaged numbers of iterations executed by the parallel algorithm for solving the test
+optimization problems}
   \label{tab:iterations}
@@
     \cline{3-8}\noalign{\smallskip}
-    \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}   \\
+    \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & &
+\multicolumn{2}{c}{$N=5$}   \\
     \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
-    \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
+    \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
+\textit{Hard}  \\
     \noalign{\smallskip}\hline
@@
 
-Если считать, что затраты на параллелизм пренебрежимо малы по сравнению с затратами на вычисление целевых функций в задачах оптимизации, то ускорение по времени от использования параллельного метода будет равно ускорению по итерациям.
-Однако, в действительности это предположение не всегда справедливо.
-Во всех численных экспериментах время вычисления целевой функции занимает примерно $10^{-3}$с.
-В таблице \ref{tab:speedup} приведено ускорение по итерациям и в куглых скобках ускорение по времени.
-В первой строке таблицы, соответствующей последовательному режиму, в скобках приведено среднее время решения одной задачи.
-Из таблицы видно, что для класов GKLS 4d выгоднее использовать один узел в многопоточном режиме, тогда как для решения более сложных пятимерных задач лучше использовать несколько узлов, каждый из которых работает в параллельном режиме.
+If one assumes the costs of parallelization to be negligible as compared to the costs of
+computing the objective functions in the optimization problems, the speedup in time due to the
+use of the parallel method would be equal to the speedup with respect to the number of
+iterations. However, actually this suggestion is not always true. In all numerical experiment the
+time of computing the objective function was approximately $10^{-3}$ sec. In
+Table~\ref{tab:speedup}, the speedups in iterations and in time (in the pendent brackets) are
+presented. In the first row of the table corresponding to the sequential mode, the averaged time
+of solving a single problem is presented in the pendent brackets. One can see from the table that
+for the GKLS 4d classes it is more efficient to utilize a single node in the multithread mode
+whereas for solving more complex five-dimensional problems, the use of several nodes is better,
+each node is operating in the parallel mode.
 
@@
 \section{Conclusions}
-В работе рассмотрены 5 различных отображений типа кривой Пеано, применяемых для редукции размерности в задачах глобальной оптимизации.
-Из предварительного сравнения, проведенного в секции \ref{sec:seq_comp}, можно сделать слудующие выводы:
+In the present work, 5 different Peano curve-type mappings applied to the dimensionality
+reduction in the global optimization problems were considered.
+From the preliminary comparison conducted in Sec.~\ref{sec:seq_comp}, one can make the
+following conclusions:
 \begin{itemize}
-  \item гладкая и неинъективная развёртки показывают наилучший результат в задачах малой размерности и могут быть успешно применены при решении задач с трудоемкими целевыми функциями.
-Особенности этих развёрток не позволяют создать на их основе масштабируемый на несколько узлов кластера алгоритм оптимизации.
-  \item сдвиговые развёртки привносят большие накладные расходы на работу метода из-за того, что требуется добавления в задачу (\ref{eq:task}) вспомогательного функционального ограничения.
-Эксперименты показали, что до 95\% поисковой информации приходится на точки, в которых вичисляется только вспомогательное ограничение.
-Сдвиговые развёртки могут быть использованы в качестве основы для параллельного алгоритма из секции \ref{sec:parallel_evolvents}, однако расходы на обработку дополнительных точек, скорее всего, приведут к малому коэффициенту ускорения от параллельности.
-Однако, при достаточно трудоёмкой целевой функции их использование может иметь смысл.
-  \item вращаемые развёртки обеспечили приемлемую скорость сходимости на задачах малой размерности в последовательном случае.
-Их использование не приводит к введению дополнительных ограничений, что позволяет построить эффективный параллельный алгоритм на их основе.
+  \item the smooth evolvent and the non-univalent one demonstrate the best result in the
+problems of small dimensionality and can be applied successfully in solving the problems with
+the computational costly objective functions. The properties of these evolvents don't allow
+developing the optimization algorithms scalable onto several cluster nodes based on these ones.
+  \item the shifted evolvents introduce large overhead costs on the operation of the method due
+to the requirement to adding an auxiliary functional constraint into the problem (\ref{eq:task}).
+The experiments have demonstrated that up to 95\% of the search information account for the
+points, in which the auxiliary constraint is computed only. The shifted evolvents can be used as
+the base for the parallel algorithm from Sec.~\ref{sec:parallel_evolvents}. However, the costs of
+processing the auxiliary points would result likely in a small speedup from the parallelization.
+However, if the objective function is computation-costly enough, the use of these evolvents
+could make sense.
+  \item the rotated evolvents have provided an acceprable speed of convergence in the problems
+of small dimensionality in the sequential mode. The use of these ones don't result in the
+introduction of the auxiliary constraints that allows constructing an efficient parallel algorithm
+based on these evolvents.
 \end{itemize}
 
-В секции \ref{sec:results_parallel} приведены результаты численных экспериментов, которые показывают, что алгоритм из раздела \ref{sec:parallel_evolvents} на основе вращаемых развёрток позволяет получить ускорение до 43 раз при решении серии задач с использованием нескольких узлов вычислитеьного кластера.
-Стоит заметить, что в расматриваемых задачах целевые функции не являются трудоёмкими (среднее время вычисления $10^{-3}$с) и в случае более сложных задач показатели ускорения по времени будут приближаться к ускорению по итерациям.
+In Sec.~\ref{sec:results_parallel} the results of the numerical experiments are presented, which
+have demonstrated the algorithm from Sec.~\ref{sec:parallel_evolvents} based on the rotated
+evolvents allowed obtaining the speedup up to 43 times when solving the problem series
+employing several nodes of the computer cluster. It is worth noting that the objective functions
+in the considered problems are not computation-costly (the averaged computation time was
+$10^{-3}$ sec). In the case of more complex problems,the speedup in time could approach the
+speedup with respect to the number of iterations.
+%\end{Russian}
 
@@
 \newblock \emph{Introduction to Global Optimization Exploiting Space-filling Curves},
-\newblock Springer, 2013.
+\newblock Springer (2013)
 
@@
 \newblock \emph{Numerical Methods in Multi-Extremal Problems (Information-Statistical Algorithms)},
-\newblock Moscow: Nauka, In Russian, 1978.
+\newblock Moscow: Nauka (1978) (In Russian)
 
@@
 \newblock R. G. Strongin,
-\newblock \emph{\emph{Algorithms for multi-extremal mathematical programming problems employing a set of joint space-filling curves}},
-\newblock \emph{J. Glob. Optim.}, \textbf{2} (1992), 357--378.
+\newblock \emph{\emph{Algorithms for multi-extremal mathematical programming problems
+employing a set of joint space-filling curves}},
+\newblock \emph{J. Glob. Optim.}, \textbf{2}, 357--378 (1992)
 
@@
 \newblock \emph{Parallel Computations for Global Optimization Problems},
-\newblock Moscow State University (In Russian), Moscow, 2013.
+\newblock Moscow State University, Moscow (2013) (In Russian)
 
-\bibitem{strSergGO}5 (MR1797058) [10.1007/978-1-4615-4677-1]
+\bibitem{strSergGO}%5 (MR1797058) [10.1007/978-1-4615-4677-1]
 \newblock R. G. Strongin and Y. D. Sergeyev,
-\newblock \emph{Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms},
-\newblock Kluwer Academic Publishers, Dordrecht (2000, 2nd ed. 2013, 3rd ed. 2014).
+\newblock \emph{Global Optimization with Non-convex Constraints. Sequential and Parallel
+Algorithms},
+\newblock Kluwer Academic Publishers, Dordrecht (2000, 2nd ed. 2013, 3rd ed. 2014)
 
@@
 \newblock \emph{Global Optimization},
-\newblock Springer, 1989.
+\newblock Springer, (1989)
+
+\bibitem{Goryachih2017}
+\newblock Goryachih, A.
+\newblock \emph{A class of smooth modification of space-filling curves for global optimization
+problems}
+\newblock Springer Proceedings in Mathematics and Statistics, \textbf{197}, pp. 57--65 (2017)
 
@@
 \newblock \emph{Theory of Global Random Search},
-\newblock Kluwer Academic Publishers, Dordrecht, 1991.
+\newblock Kluwer Academic Publishers, Dordrecht (1991)
 
 \bibitem{Strongin1991}
-Strongin, R.G.: Parallel multi-extremal optimization using a set of evolvents. Comp. Math. Math. Phys. \textbf{31(8)}, 37--46 (1991)
+Strongin, R.G.: \emph{Parallel multi-extremal optimization using a set of evolvents}. Comp. Math.
+Math. Phys. \textbf{31(8)}, 37--46 (1991)
 
 \bibitem{Gergel2009}
-Strongin, R.G., Gergel, V.P., Barkalov, K.A.: Parallel methods for global optimization problem solving. Journal of instrument engineering. \textbf{52}, 25--33 (2009) (In Russian)
+Strongin, R.G., Gergel, V.P., Barkalov, K.A.: \emph{Parallel methods for global optimization problem
+solving.} Journal of instrument engineering. \textbf{52}, 25--33 (2009) (In Russian)
 
 \bibitem{Gaviano2003}
-Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
-ACM Transactions on Mathematical Software 29(4), 469--480 (2003)
+Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: \emph{Software for generation of classes
+of test functions with known local and global minima for global optimization.} ACM
+Transactions on Mathematical Software \textbf{29(4)}, 469--480 (2003)
 
 \bibitem{grishaginClass}
-Grishagin, V.A.: Operating Characteristics of Some Global Search Algorithms. Problems of Statistical Optimization 7, 198--206 (1978) (In Russian)
+Grishagin, V.A.: \emph{Operating Characteristics of Some Global Search Algorithms.} Problems of
+Statistical Optimization \textbf{7}, 198--206 (1978) (In Russian)
 
 \bibitem{globalizerSystem}
-Gergel V.P., Barkalov K.A., and Sysoyev A.V: Globalizer: A novel supercomputer software system for solving time-consuming global optimization problems.
-Numerical Algebra, Control \& Optimization 8(1), 47--62, 2018
+Gergel V.P., Barkalov K.A., and Sysoyev A.V: Globalizer: \emph{A novel supercomputer software
+system for solving time-consuming global optimization problems.} Numerical Algebra, Control
+\& Optimization \textbf{8(1)}, 47--62 (2018)
+
+\bibitem{Sergeyev2006}
+Sergeyev, Ya.D., Kvasov, D.E.: \emph{Global search based on efficient diagonal partitions and a set
+of Lipschitz constants}. SIAM J. Optim \textbf{16(3)}, 910--937 (2006)
+
+\bibitem{SergeyevKvasov2015}
+Sergeyev, Y.D., Kvasov, D.E.: \emph{A deterministic global optimization using smooth diagonal
+auxiliary functions.} Communications in Nonlinear Science and Numerical Simulation. \textbf{21(1-3)},
+99--111 (2015)
+
+\bibitem{Zilinskas2008}
+\v Zilinskas, J.: \emph{Branch and bound with simplicial partitions for global optimization.} Math.
+Model. Anal. \textbf{13(1)}, 145--159 (2008)
+
+\bibitem{Zilinskas2014}
+Paulavi\v cius, R., \v Zilinskas, J.: \emph{Simplicial Lipschitz optimization without the Lipschitz
+constant.} J. Glob. Optim. \textbf{59(1)}, 23-40 (2014)
+
+\bibitem{Jones}
+Jones, D.R.: The direct global optimization algorithm. In: Floudas, C.A., Pardalos, P.M.
+(eds.) \emph{The Encyclopedia of Optimization}, 2nd edn., pp. 725--735. Springer, Heidelberg
+(2009)
+
+\bibitem{Gablonsky}
+Gablonsky, J.M., Kelley, C.T.: \emph{A locally-biased form of the DIRECT algorithm.} J. Glob.
+Optim., \textbf{21(1)}, 27--37 (2001)
+
+\bibitem{Evtushenko}
+Evtushenko, Y., Posypkin, M.: \emph{A deterministic approach to global box-constrained
+optimization.} Optim. Lett. \textbf{7(4)}, 819--829 (2013)
 

### 2018-05-16 2018/RSD/MultiEvolvents/RSD.tex
RSD: fixe typos

@@
 One of the possible ways to overcome the negative effects of using a numerical
-approximation of evolvent (it destroys the information about the neighbour points in $\mathbb{R}^N$ space,
+approximation of evolvent (it destroys the information about the neighbor points in $\mathbb{R}^N$ space,
 see \cite{Strongin1992}) consists in using the multiple mappings
@@
 could make sense.
-  \item the rotated evolvents have provided an acceprable speed of convergence in the problems
+  \item the rotated evolvents have provided an acceptable speed of convergence in the problems
 of small dimensionality in the sequential mode. The use of these ones don't result in the

### 2018-05-17 2018/ITNT/ITNT2018.tex
Long version

@@
 \usepackage{amssymb}
+
 \begin{document}
@@
 University of Nizhny Novgorod \cite{Strongin2000,Strongin2013} and implemented in the Globalizer 
-system \cite{Globalizer} are presented.
+system \cite{Globalizer,Globalizer1} are presented.
 
@@
 
-The parallel index algorithm, which can be applied to solving such problems 
-with partially defined functions and which has been implemented in the 
-Globalizer system, described in details in \cite{Strongin2013}. The main idea of the algorithm 
-consists in the reduction of the initial multidimensional problem to a set of 
-the nested subproblems of less dimensionality and the solving of these ones 
-in parallel. The dimensionality reduction schemes utilized in the operation 
-of the index algorithm are described briefly in the next section.
+The main idea of the index algorithm which can be applied to solving such problems 
+with partially defined functions consists in the reduction of the initial multidimensional problem to a set of 
+the subproblems of less dimensionality and the solving of these ones 
+in parallel. The index method and dimensionality reduction schemes are described briefly in the next sections.
+
+\section{Parallel index method}
+
+In this section we consider a one-dimensional multiextremal optimization problem
+\[
+\varphi(x^\ast)=\min{\left\{\varphi(x): x \in [a,b], \; g_i(x)\leq 0, \; 1 \leq i \leq m\right\}}.
+\]
+Let us describe a rules of parallel index method for the case when the algorithm carries out during one iteration $p>1$ trials simultaneously (each trial in a separate thread or process).
+Let $k(n)$ be the total number of trials, carried out after $n$ parallel iterations.
+
+Suppose $n\geq 1$  iterations of the method have been carried out, in which trials were performed at $k=k(n)$ points $x^i,1\leq i \leq k$.
+Then points $x^{k+1},\dots,x^{k+p}$  of search trials of the next $n+1$ iteration are determined according to the following rules.
+
+Rule 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the
+lower indices in ascending order of coordinate values, i.e.
+\[
+0=x_0<x_1<\dots <x_k<x_{k+1}=1,
+\]
+and juxtapose to them the values $z_i=g_\nu(x_i), \; \nu=\nu(x_i), \; 1
+\leq i \leq k$, computed at these points. The points $x_0=0$ and
+$x_{k+1}=1$ are introduced additionally, while the values $z_0$ and
+$z_{k+1}$ are not defined.
+
+Rule 2. Classify the indices $i, \; 1 \leq i \leq k$, of the trial points
+according to the number of the problem constraints fulfilled at these
+points, by constructing the sets
+\[
+I_\nu =\left\{i:1 \leq i \leq k, \; \nu=\nu(x_i) \right\}, \; 1 \leq \nu \leq m+1,
+\]
+containing the numbers of all the points $x_i, \; 1 \leq i \leq k$, with
+the same values of $\nu$. The end points $x_0=0$ and $x_{k+1}=1$ are
+interpreted as the ones having indices equal to zero. An additional set,
+$I_0=\left\{0,k+1\right\}$, corresponds to them.
+
+Determine the maximum value of the index:
+\[
+M=\max\left\{\nu(x_i), \; 1 \leq i \leq k \right \}.
+\]
+
+Rule 3. Compute the current lower estimates,
+\begin{equation}\label{Rule_Mu}
+\mu = \max\left\{ \frac{\left|z_i-z_j\right|}{ x_i - x_j }, \; i,j \in I_\nu, \; i>j \right\},
+\end{equation}
+for the unknown Lipschitz constants $L_\nu$ of the functions $g_\nu(x),1
+\leq \nu \leq m+1$. If a set $I_\nu$ contains less than two elements, or
+if $\mu_\nu$ is equal to zero, then assume $\mu_\nu=1$.
+
+Rule 4. For all nonempty sets $I_\nu, \; 1 \leq \nu \leq m+1$, compute the
+estimates
+\[
+z_\nu^\ast = \left\{
+   \begin{array}{lr}
+     -\epsilon_\nu, & \nu < M,\\
+     \min\{ g_\nu(x_i): i\in I_\nu \}, & \nu = M,
+   \end{array}
+\right.
+ \]
+where the nonnegative numbers $(\epsilon_1,...,\epsilon_m)$ are parameters
+of the algorithm.
+
+Rule 5. For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute
+the \textit{characteristics} $R(i)$ :
+\[
+R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
+\]
+\begin{equation}\label{Rule_R}
+R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
+\end{equation}
+\[
+R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
+\]
+where $\Delta_i=x_i - x_{i-1}$. The values $r_\nu > 1, \; 1 \leq
+\nu \leq m+1,$ are parameters of the algorithm. An appropriate selection
+of $r_\nu$ allows to consider the product $r_\nu \mu_\nu$ as an estimate
+of the Lipschitz constants $L_\nu, \; 1 \leq \nu \leq m+1$.
+
+Rule 6. Arrange characteristics  $R(i), 1 \leq i \leq k+1$, in decreasing order 
+\begin{equation}\label{Rule_Max}
+R(t_1)\geq R(t_2)\geq \dots \geq R(t_{k}) \geq R(t_{k+1})
+\end{equation}
+and select $p$ maximum characteristics with interval numbers $t_j, 1\leq j \leq p$.
+
+Rule 7. Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), 1\leq j\leq p$, calculated using the formulae
+\[
+x^{k+j} = \frac{x_{t_j} + x_{t_j-1}}{2}, \textrm{ if } \nu(x_{t{_j}-1}) \neq \nu(x_{t_j}),
+\]
+\begin{equation}\label{Rule_X}
+x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - \frac{z_{t_j}-z_{t_j-1}}{2r_\nu \mu_\nu}, \textrm{ if } \nu=\nu(x_{t_j-1})=\nu(x_{t_j}).
+\end{equation}
+
+The algorithm terminates if the condition $\Delta_{t_j}<\epsilon$ is satisfied at least for one number $t_j, 1 \leq j \leq p$ ; $\epsilon>0$ is the predefined accuracy.
+
+This method of parallelization has the following justification.
+The characteristics of intervals (\ref{Rule_R}) used in the index method can be considered as probability measures of the global minimum point location in these intervals.
+Inequalities (\ref{Rule_Max}) arrange intervals according to their characteristics, and trials are carried out in parallel in the first $p$ intervals with the largest probabilities.
 
@@
 condition (see \cite{Strongin2000}), i.e.
-\[
+\begin{equation}\label{holder}
 \left|g_i(y(x'))-g_i (y(x''))\right| \leq H_i \left|x'-x'' \right|^{1/N}, \; 
 x',x''\in [0,1], \; 1\leq i \leq m+1.
-\]
+\end{equation}
 Here $N$ is the dimensionality of the initial multidimensional problem and
@@
 the initial problem by the inequalities $H_i \leq 2L_i \sqrt{N+3}$.
+
+Relation (\ref{holder}) allows modification of the parallel index algorithm described in previous section for solving multidimensional problems reduced to one-dimensional ones.
+For this purpose the intervals lengths $\Delta_i$ from the rules (\ref{Rule_Mu}) and (\ref{Rule_R}) of the index method are replaced with the lengths in a new metric
+\[
+\Delta_i = \left(x_i-x_{i-1}\right)^{1/N},
+\]
+and formula (\ref{Rule_X}) is replaced by the expression
+\[
+x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - \frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t_j-1})=\nu(x_{t_j}).
+\]
+
+Various modifications of this algorithm and the corresponding theory of convergence are given in \cite{Strongin2000,Strongin2013,Gergel2005}.
 
@@
 
-In the present study, this scheme has been applied for $M=2$, i.e. only one 
-nesting level has been used
+The number of vectors $M$ and the quantity of components in each vector $N_1, N_2,...,N_M$ are the parameters of the nested optimization scheme and can be used for the forming of the subproblems with necessary properties.
+For example, if  $M=N$ i.e.
+$u_i=y_i, 1\leq i \leq N,$ each nested subproblem is a one-dimensional one.
+If $M=1$, i.e.
+$u=u_1=y$ , the solving of the problem is equivalent to solving this one using a single evolvent mapping $[0,1]$ into $D$; the nested subproblems are absent.
+
+In the present study, this scheme has been applied for $M=2$
 \[
@@
 m\right\}}= \min_{u_1\in D_1}\min_{u_2\in D_2}{\left\{\varphi(y): \; 
-g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}.
+g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
 \]
+i.e. only one nesting level has been used.
 
@@
 implemented in the Globzlizer solver developed in Lobachevsky State 
-University of Nizhny Novgorod \cite{Globalizer}. The global search methods and various 
+University of Nizhny Novgorod \cite{Globalizer,Globalizer1}. The global search methods and various 
 dimensionality reduction schemes make the algorithmic base for the 
@@
 	\item one could control the size of the feasible domain $Q_{m+1}$ with 
-respect to the whole domain $D$;
+respect to the whole search domain $D$;
 	\item the global minimizer of the objective function is known a priori 
@@
 
-%The results of experiments performed on a series of test problems of the constrained global optimization with different time of computing the the problem functions demonstrate the index global optimization algorithm combined with the block recursive scheme for the dimensionality reduction to provide a good speedup on the considered problem class.
+%The results of experiments performed on a series of test problems of the constrained global optimization with different time of computing the problem functions demonstrate the index global optimization algorithm combined with the block recursive scheme for the dimensionality reduction to provide a good speedup on the considered problem class.
 The directions of further investigations include the expansion of the number of algorithms used for solving the subproblems on the GPUs.
@@
 \section*{Acknowledgments}
-This study was supported by the Russian Science Foundation, project No 16-11-10150.
+This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods
+and software tools for the time consuming decision making problems with using supercomputers of superior performance''.
+
+\bigskip
 
@@
 \bibitem{Globalizer}
-Sysoyev A V, Barkalov K A, Sovrasov V V, Lebedev I G and Gergel V P 2017 \textit{Lecture Notes in Computer Science} \textbf{10421} 492--99
+Sysoyev A V, Barkalov K A, Sovrasov V V, Lebedev I G and Gergel V P 2017 \textit{Lect. Notes Comput. Science} \textbf{10421} 492--499
+
+\bibitem{Globalizer1}
+Gergel  V, Barkalov K and Sysoyev A 2018 \textit{Numer. Algebra Contr. Optim.} \textbf{8(1)} 47--62
 
 \bibitem{Jones}
-Floudas C A and Pardalos P M 2009 \textit{The Encyclopedia of Optimization } (Heidelberg: Springer) p 725--35
+Floudas C A and Pardalos P M 2009 \textit{The Encyclopedia of Optimization } (Heidelberg: Springer) p 725--735
 
 \bibitem{Zilinskas}
-Paulavi\v{c}ius R, \v{Z}ilinskas J and Grothey A 2011 \textit{Optim. Meth. Soft.} \textbf{26(3)} 487--98
+Paulavi\v{c}ius R, \v{Z}ilinskas J and Grothey A 2011 \textit{Optim. Meth. Soft.} \textbf{26(3)} 487--498
 
 \bibitem{Evtushenko2009}
-Evtushenko Yu G, Malkova V U and Stanevichyus A A 2009 \textit{Comput. Math. Math. Phys.} \textbf{49(2)} 246--60.
+Evtushenko Yu G, Malkova V U and Stanevichyus A A 2009 \textit{Comput. Math. Math. Phys.} \textbf{49(2)} 246--260.
 
 \bibitem{Evtushenko2014}
-Evtushenko Yu G and Posypkin M A 2014 \textit{Autom. Remote Control} \textbf{75(6)} 1025--40
+Evtushenko Yu G and Posypkin M A 2014 \textit{Autom. Remote Control} \textbf{75(6)} 1025--1040
 
 \bibitem{Strongin2017}
-Barkalov K A and Strongin R G 2017 \textit{Lecture Notes in Computer Science} \textit{10556} 18--33
+Barkalov K A and Strongin R G 2017 \textit{Lect. Notes Comput. Science} \textit{10556} 18--33
 
@@
 \bibitem{BarkalovLebedev}
-Barkalov K A and Lebedev I G 2016 \textit{Communications in Computer and Information Science} \textbf{687} 224--35
+Barkalov K A and Lebedev I G 2016 \textit{Commun. Comp. Inf. Science} \textbf{687} 224--235
+
+\bibitem{Gergel2005}
+Gergel V P and Strongin R G 2005 \textit{Lect. Notes Comput. Science} \textit{2763} 76--88
+
+\bibitem{BarkalovGergelLebedev}
+Barkalov K, Gergel V, Lebedev I 2015 \textit{Lect. Notes Comput. Science} \textbf{9251} 307--318
 
@@
 
-\bibitem{BarkalovGergelLebedev}
-Barkalov K, Gergel V, Lebedev I 2015 \textit{Lecture Notes in Computer Science} \textbf{9251} 307--18
-
 \bibitem{Gergel2017}
-Gergel V 2017 \textit{Lecture Notes in Computer Science} \textbf{10556} 314--19
+Gergel V 2017 \textit{Lect. Notes Comput. Science} \textbf{10556} 314--319
 

### 2018-05-18 2018/ITNT/ITNT2018.tex
Evolvent figures
Tables and figures numbers are corrected

@@
 numerical construction of the evolvents and the corresponding theory are 
-considered in detail in \cite{Strongin2000}.
+considered in detail in \cite{Strongin2000}. Examples of the evolvent with different $m$ in two dimensions are given in figure~\ref{fig:0}.
+
+\begin{figure}[t]
+\begin{minipage}{0.32\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1a.JPG} \\ (a)}
+\end{minipage}
+%\hfill
+\begin{minipage}{0.32\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (b)}
+\end{minipage}
+\begin{minipage}{0.32\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (c)}
+\end{minipage}
+\caption{Evolvents in two dimensions with (a) $m=3$, (b) $m=4$ and (c) $m=5$}
+\label{fig:0}
+\end{figure}
 
@@
 The level lines of the two-dimensional subproblems based on the functions 
-$f(y_1,y_2)$ and $g(y_3,y_4)$ are shown in figure~1. The feasible domains are highlighted by color. One can see that 
+$f(y_1,y_2)$ and $g(y_3,y_4)$ are shown in figure~\ref{fig:2}. The feasible domains are highlighted by color. One can see that 
 the subproblem (a) has more complex structure as compared to the subproblem 
@@
 \end{minipage}
-\center{{\bf Figure 1.} The level lines of (a) hard and (b) simple 
-subproblems}
+\caption{The level lines of (a) hard and (b) simple subproblems}
+\label{fig:2}
 \end{figure}
@@
 Let us conduct the first experiment by solving the \textit{Simple} and \textit{Hard} problem 
-series on a single node employing both available CPUs in full (i. e. $p=16$ 
-cores available). In table 1 the averaged time (in seconds) 
+series on a single node employing both available CPUs in full (i.e. $p=16$ 
+cores available). In table \ref{tab:1} the averaged time (in seconds) 
 required to solve the problems of the series is presented.
 
-\begin{table}[ht]
-\center{{\bf Table 1.} Average time for solving the problem on one node}
+\begin{table}
+\caption{Average time for solving the problem on one node}
+\label{tab:1}
 \begin{center}
@@
 graphic accelerators. Thus, at $p=32$ total 96 GPU accelerators were employed 
-each including 2688 CUDA cores. In tables 2 and 3 the average time and the speedup with respect to the run 
+each including 2688 CUDA cores. In tables \ref{tab:2} and \ref{tab:3} the average time and the speedup with respect to the run 
 on a single node are presented.
 
-\begin{table}[ht]
-\center{{\bf Table 2.} Average time for solving the problem on $p$ nodes}
+\begin{table}
+\caption{Average time for solving the problem on $p$ nodes}
+\label{tab:2}
 \begin{center}
@@
 
-\begin{table}[ht]
-\center{{\bf Table 3.} Speedup with respect to one node}
+\begin{table}
+\caption{Speedup with respect to one node}
+\label{tab:3}
 \begin{center}

### 2018-05-21 2018/PAVT_Toropov/PAVT2018_MAM.tex
FINAL

@@
 
-\documentclass[runningheads,a4paper]{llncs}
+\documentclass[a4paper]{llncs}
 \usepackage{amssymb,amsmath}
@@
 \newblock Simulation approach to structural optimization.
-\newblock Structural Optimization \textbf{1}, 37--46 (1989).
+\newblock Structural Optimization \textbf{1}, 37--46 (1989)
 
@@
 \newblock In: Sydow, A. (ed.) Proceedings of the 4th international symposium on Systems analysis and simulation,
-pp.\,207--212, Elsevier,  (1992).
+pp.\,207--212, Elsevier,  (1992)
 
@@
 \newblock Multiparameter structural optimization using FEM and multipoint explicit approximations.
-\newblock Structural Optimization \textbf{6}, 7--14 (1993).
+\newblock Structural Optimization \textbf{6}, 7--14 (1993)
 
@@
 \newblock In:  49th AIAA/ASME/ASCE/AHS/ASC Structures, Structural Dynamics, and Materials Conference, AIAA, art. no. 2008-2056
-(2008).
+(2008)
 
@@
 \newblock In: Proc. of 7th ASMO UK / ISSMO Conf. on Engineering Design Optimization. Process and Product Improvement
-(2008).
+(2008)
 
@@
 \newblock In: 13th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference
-(2010).
+(2010)
 
@@
 \newblock In: McCarthy, J.M. (ed.) Proc. of ASME Design Engineering Technical Conferences and Computers
-  in Engineering Conf., 1--12, ASME, Irvine CA (1996).
+  in Engineering Conf., 1--12, ASME, Irvine CA (1996)
 
@@
 \newblock In: 6th AIAA/NASA/ISSMO Symp. on Multidisciplinary Analysis and Optimization,  pp.\,941--951, AIAA
-(1996).
+(1996)
 
@@
 \newblock Use of mid-range approximations for optimization problems with functions of domain-dependent calculability.
-\newblock In: 3rd ISSMO/UBCAD/UB/AIAA World Congress of Structural and Multidisciplinary Optimization (1999).
+\newblock In: 3rd ISSMO/UBCAD/UB/AIAA World Congress of Structural and Multidisciplinary Optimization (1999)
 
@@
 \newblock The use of simplified numerical models as mid-range approximations.
-\newblock In: 6th AIAA/NASA/ISSMO Symp. on Multidisciplinary Analysis and Optimization, pp.\,952--958 (1996).
+\newblock In: 6th AIAA/NASA/ISSMO Symp. on Multidisciplinary Analysis and Optimization, pp.\,952--958 (1996)
 
@@
 \newblock In: Parmee, I.C. (ed.) Adaptive Computing in Design and Manufacture, 3rd Int. Conf.,
-   pp.\,21--24, PEDC, Dartington (1998).
+   pp.\,21--24, PEDC, Dartington (1998)
 
@@
 \newblock Structural and Multidisciplinary Optimization \textbf{45}(4), 515--527
-(2012).
+(2012). doi:~10.1007/s00158-011-0692-1
 
@@
 \newblock Surfaces generated by moving least squares methods.
-\newblock Mathematics of Computation \textbf{87}, 141--158 (1981).
+\newblock Mathematics of Computation \textbf{87}, 141--158 (1981)
 
@@
 \newblock An interpolation method for an irregular net of nodes.
-\newblock Int. J. Num. Meth. \textbf{20}, 1599--1612 (1984).
+\newblock Int. J. Num. Meth. \textbf{20}, 1599--1612 (1984)
 
@@
 \newblock In: 4th World Congress of Structural and Multidisciplinary Optimization, Dalian, China
-(2001).
+(2001)
 
@@
 \newblock In: Herskovits, J., Mazorche, S., Canelas, A. (eds.) 6th World Congress of Structural and Multidisciplinary Optimization, art. no. 9412
-(2005).
+(2005)
 
@@
 \newblock In: Proceedings of the 51st AIAA/ASME/ASCE/AHS/ASC Structures, Structural Dynamics, and Materials Conference, AIAA, art. no. 2010-2985
-(2010).
+(2010)
 
@@
 \newblock In: 15th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference, AIAA, art. no. 2014-2437
-(2014).
+(2014). doi:~10.2514/6.2014-2437
 
@@
 \newblock New developments in structural optimization using adaptive mesh refinement and multi-point approximations.
-\newblock Engineering Optimization \textbf{29}, 217--234 (1997).
+\newblock Engineering Optimization \textbf{29}, 217--234 (1997). doi:~10.1080/03052159708940994
 
@@
 \newblock In: 12th AIAA/ISSMO Multidisciplinary Analysis and Optimization Conference, AIAA, art. no. 2008-5803
-(2008).
+(2008).doi:~10.2514/6.2008-5803
 
@@
 \newblock Multidiscipline Design Optimization.
-\newblock Vanderplaats Research \& Development, Inc., Colorado Springs, (2001).
+\newblock Vanderplaats Research \& Development, Inc., Colorado Springs (2001)
 

### 2018-05-22 2018/Journal of Supercomputing/JoS_Paper.tex
Last version

@@
 The non-convex constraints are accounted for by a novel index scheme without utilizing the ideas of penalty functions.
-The efficiency of the parallel algorithm was confirmed experimentally by solving numerically of several hundred multidimensional multiextremal problems with non-convex constrains generated randomly.
+The efficiency of the parallel algorithm was confirmed experimentally by solving numerically several hundred multidimensional multiextremal problems with non-convex constrains generated randomly.
 \keywords{Global optimization \and Constrained problems \and Non-convex constraints \and Dimensionality reduction \and Parallel algorithms}
@@
 The objective function and constraints are assumed to satisfy the Lipschitz condition with a priori unknown Lipschitz constants.
-The analytical form of the problem’s functions may also be unknown, i.e.
+The analytical form of the problem's functions may also be unknown, i.e.
 they can be set by an algorithm computing their values at various points within the search domain (the so-called ``black-box'' functions).
@@
 To construct the set $Y^S(x)$, different approaches can be used.
-For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^i(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$.
-The set of Peano curves thus constructed allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$ for any close multidimensional images $y', y''$ from $D$, which differ only in one coordinate.
+For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^s(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$.
+The set of Peano curves thus constructed allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$, for any close multidimensional images $y', y''$ from $D$, which differ only in one coordinate.
 
-With the use of multiple mapping, it is possible to solve the initial problem (\ref{problem}) (1) by solving in parallel the problems
+With the use of multiple mapping, it is possible to solve the initial problem (\ref{problem}) by solving in parallel the problems
 \[
@@
 Note that when solving applied optimization problems, the computation of the problem function values even in one point is a computationally costly operation.
-In this case, the data transfer overheads will not affect the total computational costs predominately in this case, and the difference between the time speedup and the iteration speedup will not be so great.
+In this case, the data transfer overheads will not affect significantly the total computational costs, and the difference between the time speedup and the iteration speedup will not be so great.
 %Описание инфраструктуры
@@
 	\item to solve the initial problem directly, without the use of the penalty functions (thus eliminating the need to select the penalty coefficient and to solve a series of unconstrained problems with different penalty coefficients);
-	\item to solve the problems where the values of the problem function are not defined everywhere (for example, the objective function values are undefined outside the feasible domains);
+	\item to solve the problems where the values of the problem function are not defined everywhere (for example, the objective function values are undefined outside the feasible domain);
 	\item to use the two-level parallelization scheme with the shared and distributed memory proposed earlier for the unconstrained optimization methods.
@@
 
-%\bibitem{Evtushenko2009} Evtushenko, Yu.G., Malkova, V.U., Stanevichyus, A.A.: Parallel global optimization of functions of several variables. Comput. Math. Math. Phys. 49(2), 246--260 (2009)
-
 \bibitem{Paulavicius2011}
@@
 
-\bibitem{Sidorov2015}
-Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of computationally intensive multiextremal optimization problems.
-Lecture Notes in Computer Science.
-9251, 505--515 (2015)
-
 \bibitem{Sergeyev2013}
@@
 
-%\bibitem{Sergeyev2006} Sergeyev, Ya.D., Kvasov, D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim. 16(3), 910--937 (2006)
-
-%\bibitem{Paulavicius2014} Paulavi\v cius, R., Sergeyev, Y.D., Kvasov, D.E., \v Zilinskas, J.
-Globally-biased \textsc{Disimpl} algorithm for expensive global optimization.
-J.
-Glob.
-Optim.
-59(2), 545--567 (2014)
-
-%\bibitem{Sergeyev2015} Sergeyev, Y.D., Kvasov, D.E.: A deterministic global optimization using smooth diagonal auxiliary functions.
-Communications in Nonlinear Science and Numerical Simulation.
-21(1-3), 99–111 (2015)
-
 \bibitem{Gergel2017}
@@
 
-%\bibitem{Barkalov2016} Barkalov, K., Gergel, V.: Parallel global optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016)
+\bibitem{Sidorov2015}
+Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of computationally intensive multiextremal optimization problems.
+Lecture Notes in Computer Science.
+9251, 505--515 (2015)
 

### 2018-05-25 2018/ITNT/ITNT2018.tex
Semi-final

@@
 
-\address{National Research Lobachevsky State University of Nizhny Novgorod, Gagarin avenue 23, 
-Nizhny Novgorod, Russia, 603950}
+\address{Lobachevsky State University of Nizhni Novgorod, Gagarin avenue 23, 
+Nizhnш Novgorod, Russia, 603950}
 
@@
 investigation of the algorithm scalability has been carried out on the 
-problem class, in which the computation costs of the functions depended on 
-the iteration point. The algorithm proposed in the present work can utilize 
+problem class, in which the computation costs of the function value evaluations can be varied at different iteration points. The algorithm proposed in the present work can utilize 
 the CPUs (for solving more complex subproblems) as well as the GPUs (for 
@@
 construction of a coverage of the search domain and the computing the 
-optimized function values at all points of this coverage. The dimensionality 
+ function values at all points of this coverage. The dimensionality 
 affects the difficulty of solving the problems of considered class crucially: 
@@
 index global optimization algorithm developed in Lobachevsky State 
-University of Nizhny Novgorod \cite{Strongin2000,Strongin2013} and implemented in the Globalizer 
+University of Nizhni Novgorod \cite{Strongin2000,Strongin2013} and implemented in the Globalizer 
 system \cite{Globalizer,Globalizer1} are presented.
@@
 system for solving the global optimization problem developed in Lobachevsky 
-State University of Nizhny Novgorod.
+State University of Nizhni Novgorod.
 
@@
 \end{equation}
-with search domain
+with the search domain
 \[
@@
 iteration of the algorithm will consist in computing the values 
-$g_1(y),...,g_\nu(y),$ where the
+$g_1(y^k),...,g_\nu(y^k),$ where the
 index $\nu \leq m$ is determined by the conditions
@@
 
-Rule 3. Compute the current lower estimates,
+Rule 3. Compute the current lower estimates
 \begin{equation}\label{Rule_Mu}
-\mu = \max\left\{ \frac{\left|z_i-z_j\right|}{ x_i - x_j }, \; i,j \in I_\nu, \; i>j \right\},
+\mu_\nu = \max\left\{ \frac{\left|z_i-z_j\right|}{ x_i - x_j }, \; i,j \in I_\nu, \; i>j \right\}, \; 1 \leq \nu \leq m+1,
 \end{equation}
@@
 $u_2=(y_{N_1+1},y_{N_1+2},...,y_{N_1+N_2})$,..., $u_M=(y_{N-N_M+1},y_{N-
-N_M+2},...,y_{N})$, at that $N_1+N_2+...+N_M=N$. 
+N_M+2},...,y_{N})$, and $N_1+N_2+...+N_M=N$. 
 The subdomains $D_i, 1 \leq i \leq M$, are projections of initial search 
@@
 
-The number of vectors $M$ and the quantity of components in each vector $N_1, N_2,...,N_M$ are the parameters of the nested optimization scheme and can be used for the forming of the subproblems with necessary properties.
+The number of vectors $M$ and the quantity of components in each vector $N_1, N_2,...,N_M$ are the parameters of the nested optimization scheme and can be used to form the subproblems with necessary properties.
 For example, if  $M=N$ i.e.
@@
 optimization scheme has been considered in details for the shared/distributed 
-memory as well as for the accelerators in \cite{Sysoyev,BarkalovLebedev}. However, in these works the 
+memory as well as for the accelerators in \cite{BarkalovGergelLebedev2016,BarkalovLebedev}. However, in these works the 
 problems were considered, in which the time of the trial execution didn't 
@@
 part $g(y_1,...,y_N)$ (depending on all problem parameters), for 
-example, $\varphi(y)= f(y)g(y)$.
+example, $\varphi(y_1,...,y_N)= f(y_1,...,y_s)g(y_1,...,y_N)$.
 
@@
 computed on an accelerator, for example, GPU. For solving a problem with such 
-a structure, one can apply the parallel recursive optimization scheme with 
+a structure, one can apply the recursive optimization scheme in parallel with 
 the use CPU at the upper nesting level and GPU at the lower one.
@@
 implemented in the Globzlizer solver developed in Lobachevsky State 
-University of Nizhny Novgorod \cite{Globalizer,Globalizer1}. The global search methods and various 
+University of Nizhni Novgorod \cite{Globalizer,Globalizer1}. The global search methods and various 
 dimensionality reduction schemes make the algorithmic base for the 
 Globalizer. The numerical experiments, the results of which are presented in 
-\cite{BarkalovGergelLebedev,BarkalovGergel} demonstrate these methods, at least, are not worse than the well known 
+\cite{GergelLebedev,GergelSidorov} demonstrate these methods, at least, are not worse than the well known 
 methods applied for similar purposes, and even overcome these ones with 
@@
 graphic accelerators. Thus, at $p=32$ total 96 GPU accelerators were employed 
-each including 2688 CUDA cores. In tables \ref{tab:2} and \ref{tab:3} the average time and the speedup with respect to the run 
+each including 2688 CUDA cores; 258048 CUDA cores in all. In tables \ref{tab:2} and \ref{tab:3} the average time and the speedup with respect to the run 
 on a single node are presented.
@@
 
-Computational experiments were carried out on a high-performance cluster of 
-Lobachevsky State University of Nizhny Novgorod. The cluster node included 
-two Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM, and three NVIDIA 
-Kepler K20 GPUs (2688 CUDA cores, 6 Gb GDDR5). 
 The results of experiments demonstrate that the parallel index algorithm combined with the nested optimization scheme provides a good speedup on the considered problem class.
@@
 
-\bibitem{Sysoyev}
-Sysoyev A V, Barkalov K A, Gergel V P and Lebedev I G 2015 \textit{CEUR Workshop Proceedings} \textbf{1482} 61--68
+\bibitem{BarkalovGergelLebedev2016}
+Barkalov K, Gergel V and Lebedev I  2016 \textit{AIP Conf. Proc.} \textbf{1738} 400006
 
@@
 \bibitem{Gergel2005}
-Gergel V P and Strongin R G 2005 \textit{Lect. Notes Comput. Science} \textit{2763} 76--88
+Gergel V P and Strongin R G 2003 \textit{Lect. Notes Comput. Science} \textit{2763} 76--88
 
-\bibitem{BarkalovGergelLebedev}
-Barkalov K, Gergel V, Lebedev I 2015 \textit{Lect. Notes Comput. Science} \textbf{9251} 307--318
+\bibitem{GergelLebedev}
+Gergel V and Lebedev I 2015 \textit{Proc. Comput. Science} \textbf{66} 53--62
 
-\bibitem{BarkalovGergel}
-Barkalov K and Gergel V 2016 \textit{J. Glob. Optim.} \textbf{66(1)} 3--20
+\bibitem{GergelSidorov}
+Gergel V and Sidorov S 2015 \textit{Lect. Notes Comput. Science} \textbf{9251} 505--515
 

### 2018-05-27 2018/RSD/MultiEvolvents/RSD.tex
RSD: fix review comments

@@
 of a multidimensional problem to solving the one-dimensional subproblems implicitly (see, for
-example, the methods of diagonal partitions \cite{Sergeyev2006,SergeyevKvasov2015} or
-simplicial partitions \cite{Zilinskas2008,Zilinskas2014}). In the present work, we will use the
+example, the methods of diagonal partitions \cite{SergeyevKvasov2015} or
+simplicial partitions \cite{Zilinskas2014}). In the present work, we will use the
 approach developed in Lobachevsky State University of Nizhni Novgorod based on the idea of
@@
 described algorithm are given, for example, in \cite{strSergGO}.
+
+The numerical experiments, the results of which are presented in \cite{GergelLebedev,GergelSidorov} demonstrate that the method, at least, is not worse than the well-known global optimization algorithms DIRECT \cite{Jones} and DIRECT\textit{l} \cite{Gablonsky}, and even overcome these ones with respect to some parameters.
 
@@
 
-\bibitem{Sergeyev2006}
-Sergeyev, Ya.D., Kvasov, D.E.: \emph{Global search based on efficient diagonal partitions and a set
-of Lipschitz constants}. SIAM J. Optim \textbf{16(3)}, 910--937 (2006)
+%\bibitem{Sergeyev2006}
+%Sergeyev, Ya.D., Kvasov, D.E.: \emph{Global search based on efficient diagonal partitions and a set
+%of Lipschitz constants}. SIAM J. Optim \textbf{16(3)}, 910--937 (2006)
 
@@
 
-\bibitem{Zilinskas2008}
-\v Zilinskas, J.: \emph{Branch and bound with simplicial partitions for global optimization.} Math.
-Model. Anal. \textbf{13(1)}, 145--159 (2008)
+%\bibitem{Zilinskas2008}
+%\v Zilinskas, J.: \emph{Branch and bound with simplicial partitions for global optimization.} Math.
+%Model. Anal. \textbf{13(1)}, 145--159 (2008)
 
@@
 Paulavi\v cius, R., \v Zilinskas, J.: \emph{Simplicial Lipschitz optimization without the Lipschitz
-constant.} J. Glob. Optim. \textbf{59(1)}, 23-40 (2014)
+constant.} J. Glob. Optim. \textbf{59(1)}, 23--40 (2014)
 
@@
 
+\bibitem{GergelLebedev}
+Gergel V. and Lebedev I. \emph{Heterogeneous Parallel Computations for Solving Global Optimization Problems.} Proc. Comput. Science \textbf{66}, pp. 53--62 (2015)
+
+\bibitem{GergelSidorov}
+Gergel V., Sidorov S.
+\emph{A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems.} {Lect.
+Notes Comput.
+Science} \textbf{9251}, pp.
+505--515 (2015)
+
 \end{thebibliography}

### 2018-06-10 2018/YSC/slides/slides.tex
YSC: added a template for slides

--- 2017-10-17 2017/URAL-PDC/slides/slides.tex
@@
 
-\title{Parallel Multi-objective Optimization Method for Finding Complete Set of Weakly Efficient Solutions}
+\title{Comparison of dimensionality reduction schemes for derivative-free global optimization algorithms}
 \author{\textbf{Vladislav~Sovrasov}}
-\institute{Lobachevsky University}
+\institute{Lobachevsky State University of Nizhni Novgorod}
 \date{}
@@
 
-  \begin{figure}[ht]
-    \subfloat{{\includegraphics[width=.35\textwidth]{img/peano3d.png} }}
-    \subfloat{{\includegraphics[width=.35\textwidth]{img/peano2d.png} }}
-  \end{figure}
+  %\begin{figure}[ht]
+  %  \subfloat{{\includegraphics[width=.35\textwidth]{img/peano3d.png} }}
+  %  \subfloat{{\includegraphics[width=.35\textwidth]{img/peano2d.png} }}
+  %\end{figure}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Scalarization tecnique}
-  Multi-objective problem can be reduced to a scalar optimization problem using the following function:
-  \begin{eqnarray*}
-    h(x,y)=\min\{f_i(x)-f_i(y):1\leqslant i\leqslant m\}, \\
-    \varphi(x)=\max\{h(x,y):y\in [0,1]\},x\in [0,1].
-  \end{eqnarray*}
-If \(x^*\) is weakly effective point, then \(\varphi(x^*)\le 0\), in the opposite case \(\varphi(x^*)> 0\), so
-the equivalent scalar problem is:
-  \begin{displaymath}
-    \varphi^*=\min\{\varphi(x):x\in [0,1]\}.
-  \end{displaymath}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Scalarization tecnique}
-  One-dimensional example:
-  \begin{figure}
-    \subfloat{{\includegraphics[width=.5\textwidth]{img/objectives.pdf} }}
-    \subfloat{{\includegraphics[width=.5\textwidth]{img/phi_func.pdf} }}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Optimization method}
-Optimization method generates searh sequence \(\{x_k\}\) and consists of the following steps:
-\begin{enumerate}
-  \setlength{\itemindent}{.1in}
-  \item[Step 1.] Sort the searh information (one-dimentional points) in increasing order.
-  \item[Step 2.] Compute an approximation of the function \(\varphi(x)\).
-  \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
-  \item[Step 4.] Choose an interval \((x_{t-1}, x_t)\) with the greatest characteristic and
-  compute objective \(f(y(x))\) in the point choosen using the descision rule \(d\):
-  \begin{displaymath}
-    x^{k+1}=d(t)\in (x_{t-1}, x_t)
-  \end{displaymath}
-  \item[Step 4.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
-\end{enumerate}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Improved parallel optimization method}
+  \frametitle{Conclusion and future work}
+    Already done:
     \begin{itemize}
-    \item
-    \textit{Local refinement.} Every \(q\) iterations ignore the characteristics \(R(i)\)
-    and perform calculation of the objective at the minimum point of the \(\varphi(x)\) approximation.
-    \item
-    \textit{Parallelization by characteristics.} At the Step 4 choose \(p\) best intervals
-    and generate \(p\) new points using rule \(d(i)\). Then compute the objective \(f(y(x))\) at this
-    point in parallel exploiting \(p\) computation units.
-
-    \enspace
-    In the best case (method doesn't generate redundand points compared to sequental
-    one and wastes time only on computation of the objective)
-    this scheme can give speedup in \(p\) times.
     \end{itemize}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results}
-  Examples of numerical solutions:
-  \begin{figure}[ht]
-      \centering
-      \subfloat{{\includegraphics[width=.5\textwidth]{../img/viennet.png}}}
-      \subfloat{{\includegraphics[width=.5\textwidth]{../img/poloni.pdf} }}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results}
-
-  Number of iterations decreases when increasing number of threads \(p\).
-  \begin{table}
-    \centering
-    \begin{tabular}{|l|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
-  \hline
-  \textbf{Problem} & \multicolumn{5}{c|}{\(p\)}\\
-  \cline{2-6}
-    & \(1\) & \(2\) & \(4\) & \(8\) & \(16\)\\
-  \hline
-  Markin-Strongin & 1041(198) & 516(198) & 256(185) & 131(197) & 68(191) \\
-  \hline
-  Fonseca and Fleming 2d & 1181(93) & 636(99) & 386(111) & 176(95) & 106(97) \\
-  \hline
-  Fonseca and Fleming 3d & 5346(160) & 3551(183) & 1186(143) & 606(153) & 351(142) \\
-  \hline
-  Viennet problem & 4896(276) & 2156(273) & 1226(270) & 631(287) & 286(274)\\
-  \hline
-  Poloni's function & 3351(102) & 1706(90) & 856(88) & 426(96) & 201(99) \\
-  \hline
-  \end{tabular}
-  \end{table}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results (speedup in iterations)}
-  \begin{table}
-    \centering
-    \begin{tabular}{|l|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
-  \hline
-  \textbf{Problem} & \multicolumn{4}{c|}{\(p\)}\\
-  \cline{2-5}
-    & \(2\) & \(4\) & \(8\) & \(16\)\\
-  \hline
-  Markin-Strongin & 2.02 & 4.07 & 7.95 & 15.31 \\
-  \hline
-  Fonseca and Fleming 2d & 1.86 & 3.06 & 6.71 & 11.14 \\
-  \hline
-  Fonseca and Fleming 3d & 1.51 & 4.51 & 8.82 & 15.23 \\
-  \hline
-  Viennet problem & 2.27 & 3.99 & 7.76 & 17.12\\
-  \hline
-  Poloni's problem & 1.96 & 3.91 & 7.87 & 16.67 \\
-  \hline
-  \end{tabular}
-  \end{table}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results (speedup in time)}
-  \begin{table}[ht]
-    \centering
-    \begin{tabular}{|l|p{1.6cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
-  \hline
-  \textbf{Problem} & \multicolumn{5}{c|}{\(p\)}\\
-  \cline{2-6}
-  &\(1\)(time, s) & \(2\) & \(4\) & \(8\) & \(16\)\\
-  \hline
-  Markin-Strongin & 104.47 & 1.97 & 3.65 & 6.79 & 9.90 \\
-  \hline
-  Fonseca and Fleming 2d & 118.95 & 1.85 & 2.81 & 5.79 & 6.40 \\
-  \hline
-  Fonseca and Fleming 3d & 554.45 & 1.51 & 4.14 & 8.05 & 10.69 \\
-  \hline
-  Viennet problem & 1488.6 & 2.22 & 3.64 & 6.98 & 13.49\\
-  \hline
-  Poloni's problem & 336.74 & 1.82 & 3.64 & 6.98 & 10.60 \\
-  \hline
-  \end{tabular}
-  \end{table}
+    Future work:
+    \begin{itemize}
+    \end{itemize}
 \end{frame}

### 2018-06-14 2018/Journal of Supercomputing/JoS_Paper.tex
Some new text has been included

@@
 %\documentclass[twocolumn]{svjour3}          % twocolumn
-%
+
+\usepackage[utf8x]{inputenc}
+\usepackage[english,russian]{babel}
+
 \smartqed  % flush right qed marks, e.g. at end of proof
@@
 It is also supposed that even a single computation of the function value can be time-consuming, since it requires performing numerical simulation in applied problems.
-%These assumptions are typical for many approaches to building parallel algorithms for unconstrained global optimization \cite{Jones2001,Paulavicius2011,Evtushenko2013}.
 
@@
 
-Among the examples of such problems are the optimal control problems described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2011,Balandin2017}.
+\Russian
+Данная ситуация является типичной для многих задач оптимального проектирования, когда при нарушении некоторых условий функционирования моделируемого объекта другие его характеристики оказываются неопределенными.
+
+Among the examples of such problems are the optimal control problems described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}.
 It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e.
@@
 
+\Russian
+В статье \cite{Barkalov2017} приведено сравнение двух разных способа учета ограничений - классического метода штрафных функций и индексной схемы.
+Экспериментально подтрвеждено превосходство индексного метода над методом штрафных функций как по числу итераций, так и по числу вычсислений ограничений и целевой функции.
+Для решения задачи с заданной точностью индексному методу требовалось в 3-5 раз меньше итераций (в зависимости от размерности задачи и доли допусимой области).
+Одновременно с этим значение целевой функции вычислялось в среднем лишь на каждой пятой итерации индексного метода, в отличие от метода штрафных функций, в котором значения целевой функции вычисляются на каждой итерации.
+
 \section{Parallel Index Algorithm with a Set of Evolvents}
@@
 The set of Peano curves thus constructed allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$, for any close multidimensional images $y', y''$ from $D$, which differ only in one coordinate.
+
+Another scheme of building of the multiple mapping was proposed  in \cite{6_Gergel2009}.
+The building of a set of Peano curves not by the shift along the main diagonal of the hypercube but by rotation of the evolvents around the coordinate origin is a distinctive feature of the proposed scheme.
+In the initial non-rotated mapping for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other.
+In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer.
+The maximum number of various rotations of the evolvents mapping the $N$-dimensional hypercube onto a one-dimensional interval is $2^N$.
+The employment of all possible rotations might appear to be redundant.
+In this case, we can select only a part of all rotations.
+As a possible approach, we propose to generate the new evolvents by the rotation of the initial evolvent on the angle of $\pm\pi/2$ in each of the coordinate planes.
+Taking into account the initial evolvent the total number of such evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals equals to $N(N-1)+1$.
+\Russian
+Именно такой подход к построению multiple mapping был использован в данной статье при проведении вычислительных экспериментов
 
@@
 
-%At every iteration of the method, the process with the index $s, 0\leq s < S,$ performs $p$ trials in parallel at the points $x^{s+iS}, 0\leq i < p$.
-At that, each process stores all $S\cdot p$ points, and an attribute indicating whether this point is blocked by another process or not is stored for each point.
-Let us remind that the point is blocked if the process starts the execution of a trial at this point.
-
-%At every iteration of the algorithm, operating within the $s$-th process, determines the coordinates of $p$ ``its own'' trial points.
-Then, the interchange of the coordinates of images of the trial points $y^{s+iS},\; 0\leq i<p,\; 0\leq s< S$ is performed (from each process to each one).
-After that, the preimages $x^{q+iS},\; 0\leq q<S,\; q\neq s$ of the points received by the $s$-th process from the neighbor ones are determined with the use of the evolvent $y^s(x)$.
-The points blocked within the $s$-th process will correspond to the preimages obtained.
-Then, each process performs the trials at the non-blocked points, the computations are performed in parallel using OpenMP.
-The results of the executed trials (the index of the point, the computed values of the problem functions, and the attribute of unblocking of this point) are transferred to all rest processes.
-All the points are added to the search information database, and the transition to the next iteration is performed.
-
 \section{Results of Numerical Experiments}
@@
 
-%GKLS generator for the functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.) has been proposed in \cite{Gaviano2003}.
-Four GKLS classes of differentiable test functions of the dimensions $N = 4$ and $5$, have been used.
-For each dimension, both \textit{Hard} and \textit{Simple} classes have been considered.
-The difficulty of a class was increased either by decreasing the radius of the attraction region of the global minimizer, or by decreasing the distance from the global minimizer $y^*$ to the domain boundaries.
-Application of the generator for studying some optimization algorithms has been described in \cite{Paulavicius2014,Sergeyev2015}.
-
 In this study, we will use a GKLS generator of the functions of arbitrary dimensionality proposed in \cite{Gaviano2003} to produce constrained problems.
 The scheme for generating constrained global optimization problems is proposed in \cite{Gergel2017}.
-Previous investigations have confirmed experimentally that the index method is not inferior to well-known analogues.
+\Russian
+Тестовые задачи с ограничениями, полученные с помощью данной схемы, обладают важным свойством, присущим прикладным задачам оптимизации: условный глобальный минимум находится на границе допустимой области, в то время как глобальный минимум целевой функции без ограничений находится вне допустимой области.
+При этом можно варьировать объем допустимой области, а также число ограничений, активных в точке минимума.
+
+%Previous investigations have confirmed experimentally that the index method is not inferior to well-known analogues.
 A comparison of the method with the well known DIRECT method \cite{Jones2001} in solving unconstrained optimization problems has been performed in \cite{Barkalov2015}.
@@
 
-The experiments have been carried out by solving a series of $100$ problems with two constraints and the objective functions from the \textit{Simple} and \textit{Hard} GKLS classes with the dimensionalities $N=4$, $N=5$.
+The experiments have been carried out by solving a series of $100$ problems with two constraints and one objective function from the \textit{Simple} and \textit{Hard} GKLS classes with the dimensionalities $N=4$, $N=5$.
 The number of the used cluster nodes $S$ and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at each node have been varied.
@@
 The size of the vicinity was selected as $\delta=0.03\left\|b-a\right\|$, where $a$ and $b$ are the boundaries of the search domain.
-For the purpose of simulation of the computational complexity inherent to applied optimization problems, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (a series summation of $80$ thousand elements).
+For the purpose of simulation of the computational complexity inherent to applied optimization problems, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (a series summation of $100$ thousand elements).
 
@@
 8(16)	&	31	&	77.7	&	&	264.6	&	374	\\
+12(1)	&	99.3	&	141.8	&	&	700.1	&	897.3	\\
+12(16)	&	31	&	77.7	&	&	264.6	&	374	\\
 		\noalign{\smallskip}\hline
@@
 \end{table}
+
+%\begin{table}
+	%\caption{Average number of iterations}
+	%\label{tab:2}
+	%\center
+	%\begin{tabular}{cccccc}
+		%\hline\noalign{\smallskip}
+		%\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
+		%\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
+		 %& \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
+		%\noalign{\smallskip} \hline \noalign{\smallskip}
+%1(1)	&	58320	&	84546	&	&	266943	&	287102	\\
+%1(16)	&	4297	&	6601	&	&	22655	&	56754	\\
+%2(1)	&	34791	&	52126	&	&	188465	&	241369	\\
+%2(16)	&	2029	&	3239	&	&	16689	&	40763	\\
+%4(1)	&	22223	&	47771	&	&	135734	&	180489	\\
+%4(16)	&	1281	&	2483	&	&	9241	&	35024	\\
+%8(1)	&	13844	&	18933	&	&	77748	&	94563	\\
+%8(16)	&	608	&	1473	&	&	5820	&	23033	\\
+%12(1)	&	13844	&	18933	&	&	77748	&	94563	\\
+%12(16)	&	608	&	1473	&	&	5820	&	23033	\\
+		%\noalign{\smallskip}\hline
+	%\end{tabular}
+%\end{table}
 
@@
 	\center
-	\begin{tabular}{cccccc}
+	\begin{tabular}{ccccccccc}
 		\hline\noalign{\smallskip}
-		\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
-		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
-		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
+		\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$}& & \multicolumn{2}{c}{$N=6$} \\
+		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \noalign{\smallskip}
+		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
 		\noalign{\smallskip} \hline \noalign{\smallskip}
-1(1)	&	58320	&	84546	&	&	266943	&	287102	\\
-1(16)	&	4297	&	6601	&	&	22655	&	56754	\\
-2(1)	&	34791	&	52126	&	&	188465	&	241369	\\
-2(16)	&	2029	&	3239	&	&	16689	&	40763	\\
-4(1)	&	22223	&	47771	&	&	135734	&	180489	\\
-4(16)	&	1281	&	2483	&	&	9241	&	35024	\\
-8(1)	&	13844	&	18933	&	&	77748	&	94563	\\
-8(16)	&	608	&	1473	&	&	5820	&	23033	\\
+1(1)	&	58320	&	84546	&	&	266943	&	287102 & & 1 & 2	\\
+1(16)	&	4297	&	6601	&	&	22655	&	56754	& &  & \\
+2(1)	&	34791	&	52126	&	&	188465	&	241369	& &  &\\
+2(16)	&	2029	&	3239	&	&	16689	&	40763	& &  &\\
+4(1)	&	22223	&	47771	&	&	135734	&	180489 & &  &	\\
+4(16)	&	1281	&	2483	&	&	9241	&	35024	& &  &\\
+8(1)	&	13844	&	18933	&	&	77748	&	94563	& &  &\\
+8(16)	&	608	&	1473	&	&	5820	&	23033	& &  &\\
+12(1)	&	13844	&	18933	&	&	77748	&	94563	& &  &\\
+12(16)	&	608	&	1473	&	&	5820	&	23033	& &  &\\
 		\noalign{\smallskip}\hline
@@
 In this case, the data transfer overheads will not affect significantly the total computational costs, and the difference between the time speedup and the iteration speedup will not be so great.
-%Описание инфраструктуры
 
@@
 This was confirmed by the results of solving numerically of several hundred test problems using 128 cores of the Lobachevsky supercomputer.
+\Russian
+Что-то про применение алгоритма. 
 %The direction of further research is the generalization of the considered parallel algorithm for solving multicriteria problems.
@@
 
-\bibitem{Balandin2011}
-Balandin, D.V., Kogan, M.M.: Optimal linear-quadratic control: From matrix equations to linear matrix inequalities. Automation and Remote Control. 72(11), 2276--2284 (2011)
+%\bibitem{Balandin2011} Balandin, D.V., Kogan, M.M.: Optimal linear-quadratic control: From matrix equations to linear matrix inequalities. Automation and Remote Control. 72(11), 2276--2284 (2011)
 
@@
 \bibitem{Gergel2017}
-Gergel, V.: An approach for generating test problems of constrained global optimization. Lecture Notes in Computer Science. 10556, 314--319 (2017)
+Gergel, V.: An approach for generating test problems of constrained global optimization. LNCS. 10556, 314--319 (2017)
 
-\bibitem{Barkalov2015}
-Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi coprocessor for solving global optimization problems. Lecture Notes in Computer Science. 9251, 307--318 (2015)
+%\bibitem{Barkalov2015} Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi coprocessor for solving global optimization problems. LNCS. 9251, 307--318 (2015)
 
 \bibitem{Sidorov2015}
-Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of computationally intensive multiextremal optimization problems.
-Lecture Notes in Computer Science.
-9251, 505--515 (2015)
+Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of computationally intensive multiextremal optimization problems. LNCS. 9251, 505--515 (2015)
+
+\bibitem{Barkalov2017}
+Barkalov, K., Lebeled, I.: Comparing Two Approaches for Solving Constrained Global Optimization Problems. LNCS. 10556, 301--306 (2017)
+
+\bibitem{6_Gergel2009}
+Strongin, R.G., Gergel, V.P., Barkalov, K.A.: Parallel methods for global optimization problem solving. Journal of instrument engineering. 52, 25--33 (2009) (In Russian)
 

### 2018-06-17 2018/Journal of Supercomputing/JoS_Paper.tex
Similarity correction

@@
 Данная ситуация является типичной для многих задач оптимального проектирования, когда при нарушении некоторых условий функционирования моделируемого объекта другие его характеристики оказываются неопределенными.
+Например, моделируемым объектом является некоторое электронное устройство, и минимизируется время его срабатывания.
+Однако время срабатывания является неопределенным, если при данном сочетании параметров устройство вообще не работает.
 
-Among the examples of such problems are the optimal control problems described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}.
+%Another example is the optimal control problem described through system of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}.
 It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e.
@@
 
-Let us consider the $N$-dimensional global optimization problem
+A constrained global optimization problem can be formulated as follows
+%Let us consider the $N$-dimensional global optimization problem
 \begin{equation}\label{problem}
@@
 \end{equation}
-The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
+The objective function $\varphi(y)$ (hereafter denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
 \[
@@
 \]
-Algorithms for numerical construction of Peano curve approximation (\textit{evolvent}) are given in \cite{Sergeyev2013}.
-Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined and computable in the domains
+Algorithms for numerical construction of Peano curve approximation (\textit{evolvent}) are considered in \cite{Sergeyev2013}.
+Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined and computable in the subranges
 \[
@@
 \end{itemize}
-The pair of values $ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} $ is a \textit{trial result}.
+The dyad $ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} $ will be referred to as the \textit{trial outcome}.
 
@@
 \Russian
-В статье \cite{Barkalov2017} приведено сравнение двух разных способа учета ограничений - классического метода штрафных функций и индексной схемы.
+Сравнение двух разных способа учета ограничений - классического метода штрафных функций и индексной схемы - приведено в статье \cite{Barkalov2017}.
 Экспериментально подтрвеждено превосходство индексного метода над методом штрафных функций как по числу итераций, так и по числу вычсислений ограничений и целевой функции.
@@
 However, a partial loss of information on the nearness of the points in the multidimensional space occurs in this case since a point $x \in [0,1]$ has only the left and the right neighbors while the corresponding point $y(x) \in R^N$ has the neighbors in $2N$ directions.
-As a result, when using the mappings like a Peano curve, the images $y', y''$, which are close to each other in the $N$-dimensional space can correspond to the preimages $x', x''$, which can be far away from each other in the interval $[0,1]$.
+As a result, when using the mappings like a Peano curve, the close images $y', y'' \in D$ can correspond to the remote preimages $x', x'' \in [0,1]$.
 This property results in excess computations since several limit points $x', x''$ of the trial sequence generated by the index method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional space.
@@
 To construct the set $Y^S(x)$, different approaches can be used.
-For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^s(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$.
+%For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^s(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$.
 The set of Peano curves thus constructed allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$, for any close multidimensional images $y', y''$ from $D$, which differ only in one coordinate.
 
-Another scheme of building of the multiple mapping was proposed  in \cite{6_Gergel2009}.
-The building of a set of Peano curves not by the shift along the main diagonal of the hypercube but by rotation of the evolvents around the coordinate origin is a distinctive feature of the proposed scheme.
+%Проверить
+For example, a scheme of building of the multiple mapping was proposed  in \cite{6_Gergel2009}.
+The building of a set of Peano curves by rotation of the evolvents around the coordinate origin is a distinctive feature of this scheme.
 In the initial non-rotated mapping for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other.
@@
 Taking into account the initial evolvent the total number of such evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals equals to $N(N-1)+1$.
+This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.
 \Russian
-Именно такой подход к построению multiple mapping был использован в данной статье при проведении вычислительных экспериментов
+Именно такой подход к построению multiple mapping был использован в данной статье при проведении вычислительных экспериментов.
 
@@
 Each one-dimensional problem is solved on a separate processor/node.
-The trial results at the point $x^k$ obtained for the problem being solved by particular processor are interpreted as the results of the trials in the remaining problems (at the corresponding points $x^{k_1},...,x^{k_S}$).
+The trial outcome at the point $x^k$ obtained for the problem being solved by particular processor are interpreted as the outcomes of the trials in the remaining problems (at the corresponding points $x^{k_1},...,x^{k_S}$).
 In this approach, a trial at the point $x^k \in [0,1]$ executed in the framework of the $s$-th problem, consists of the following sequence of operations.
@@
 \]
-is the result of the trial at the point $x^k$.
+is the outcome of the trial at the point $x^k$.
 
-\item Determine the preimages $x^{k_s} \in [0,1]$, $1 \leq s \leq S$, of the point $y^k$ and interpret the trial executed at the point $y^k \in D$ as the execution of the trials at the $S$ points  $x^{k_1},...,x^{k_S}$ with the same results
+\item Determine the preimages $x^{k_s} \in [0,1]$, $1 \leq s \leq S$, of the point $y^k$ and interpret the trial executed at the point $y^k \in D$ as the execution of the trials at the $S$ points  $x^{k_1},...,x^{k_S}$ with the same outcomes
 \[
@@
 
-\item Inform the rest of processors about the trial results at the point $y^k$.
+\item Inform the rest of processors about the trial outcome at the point $y^k$.
 
@@
 Тестовые задачи с ограничениями, полученные с помощью данной схемы, обладают важным свойством, присущим прикладным задачам оптимизации: условный глобальный минимум находится на границе допустимой области, в то время как глобальный минимум целевой функции без ограничений находится вне допустимой области.
-При этом можно варьировать объем допустимой области, а также число ограничений, активных в точке минимума.
+При этом можно задавать объем допустимой области и  число ограничений, активных в точке минимума, тем самым усложняя или упрощая постановку задачи.
 
@@
 The number of the used cluster nodes $S$ and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at each node have been varied.
-The problem was considered to be solved, if the algorithm generated a trial point $y^k$ in the $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^*\right\|\leq \delta$.
-The size of the vicinity was selected as $\delta=0.03\left\|b-a\right\|$, where $a$ and $b$ are the boundaries of the search domain.
+The problem was considered to be solved, if the algorithm generated a trial point $y^k$ in the $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^*\right\|\leq \delta = 0.03\left\|b-a\right\|$.
 For the purpose of simulation of the computational complexity inherent to applied optimization problems, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (a series summation of $100$ thousand elements).
 
-Computational experiments were carried out on a high-performance cluster of Lobachevsky State University of Nizhni Novgorod.
-The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.
+Computational experiments were carried out on Lobachevsky supercomputer.
+The node of supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.
 The CPU had 8 cores, i.e.
 each node had a total of 16 cores.
-
 The average time and the number of iterations, which were required to solve the problems of the series at various parallelization parameters are given in Tables 1 and 2.
@@
 
-The results demonstrate the speedup achieved when using the shared memory at a node (performing several trials within a problem in parallel) as well as the distributed memory (solving in parallel several subproblems at different nodes).
+The results demonstrate the speedup achieved when using the shared memory at a node as well as the distributed memory.
 In this case, the highest time speedup was 10 (when using 64 cores on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster nodes).
@@
 
-The parallel index method for solving constrained global optimization problems considered in this paper offers the following possibilities:
+\Russian
+В заключение отметим, что подавляющее большинство оптимизационных алгоритмов (за исключением методов, основанных на идеях случайного поиска) реализуют стратегию целенаправленного выбора.
+Указанная стратегия состоит в том, чтобы на основе анализа небольшой части вариантов исключить из дальнейшего рассмотрения многие заведомо не перспективные случаи и сосредоточить дальнейший поиск в подмножестве, содержащем лучший вариант.
+
+Сложность моделируемых процессов и объектов может приводить к тому, что их характеристики не будут обладать свойством монотонности, что обуславливает недостаточность методов локального поиска для оценки наилучшего варианта.
+В таких задачах должны применяться методы глобального поиска, которые обеспечивают целенаправленность поиска за счет учета ограниченности изменения характеристик моделируемого объекта при ограниченных изменениях его параметров.
+Математическая формулировка этого факта может иметь форму Lipschitz condition, uniform H\"older condition и т.п.
+Именно к таким методам, способным решать сложные прикладные задачи оптимизации, относится рассмотренный в данной статье параллельный индексный алгоритм.
+
+The parallel index method for solving constrained global optimization problems %considered in this paper 
+offers the following possibilities:
 \begin{itemize}
@@
 This was confirmed by the results of solving numerically of several hundred test problems using 128 cores of the Lobachevsky supercomputer.
+
 \Russian
-Что-то про применение алгоритма. 
-%The direction of further research is the generalization of the considered parallel algorithm for solving multicriteria problems.
+Перспективным направлением дальнейшей работы будет являться применение множественных отображений для редукции размерности в задачах многокритериальной оптимизации.
+В подобных задачах построение множества Парето требует решения целой серии скалярных задач.
+При их параллельном решении можно ожидать более высокого ускорения, чем при решении одной скалярной задачи.
 
@@
 
-\bibitem{Balandin2017}
-Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 (2017)
+%\bibitem{Balandin2017}Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 (2017)
 

### 2018-06-21 2018/Journal of Supercomputing/JoS_Paper.tex
Some modifications in Russian

@@
 Сравнение двух разных способа учета ограничений - классического метода штрафных функций и индексной схемы - приведено в статье \cite{Barkalov2017}.
-Экспериментально подтрвеждено превосходство индексного метода над методом штрафных функций как по числу итераций, так и по числу вычсислений ограничений и целевой функции.
+В ней экспериментально подтрвеждено превосходство индексного метода над методом штрафных функций как по числу итераций, так и по числу вычсислений ограничений и целевой функции.
 Для решения задачи с заданной точностью индексному методу требовалось в 3-5 раз меньше итераций (в зависимости от размерности задачи и доли допусимой области).
@@
 For example, a scheme of building of the multiple mapping was proposed  in \cite{6_Gergel2009}.
-The building of a set of Peano curves by rotation of the evolvents around the coordinate origin is a distinctive feature of this scheme.
-In the initial non-rotated mapping for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other.
+The building of a Peano curves set by rotation of the evolvents around the coordinate origin is a distinctive feature of this scheme.
+In the initial non-rotated evolvent for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other.
 In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer.
@@
 \Russian
-Именно такой подход к построению multiple mapping был использован в данной статье при проведении вычислительных экспериментов.
+Именно такой подход к построению set of evolvents был использован в данной статье при проведении вычислительных экспериментов.
 
@@
 
-%Previous investigations have confirmed experimentally that the index method is not inferior to well-known analogues.
-A comparison of the method with the well known DIRECT method \cite{Jones2001} in solving unconstrained optimization problems has been performed in \cite{Barkalov2015}.
 In the present study, an experimental investigation is carried out of the speedup, which is obtained by the use of the index method in combination with the two-level parallelization scheme from \cite{Sidorov2015}.
@@
 8(16)	&	31	&	77.7	&	&	264.6	&	374	\\
-12(1)	&	99.3	&	141.8	&	&	700.1	&	897.3	\\
-12(16)	&	31	&	77.7	&	&	264.6	&	374	\\
+10(1)	&	99.3	&	141.8	&	&	700.1	&	897.3	\\
+10(16)	&	31	&	77.7	&	&	264.6	&	374	\\
 		\noalign{\smallskip}\hline
@@
 \end{table}
-
-%\begin{table}
-	%\caption{Average number of iterations}
-	%\label{tab:2}
-	%\center
-	%\begin{tabular}{cccccc}
-		%\hline\noalign{\smallskip}
-		%\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
-		%\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
-		 %& \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
-		%\noalign{\smallskip} \hline \noalign{\smallskip}
-%1(1)	&	58320	&	84546	&	&	266943	&	287102	\\
-%1(16)	&	4297	&	6601	&	&	22655	&	56754	\\
-%2(1)	&	34791	&	52126	&	&	188465	&	241369	\\
-%2(16)	&	2029	&	3239	&	&	16689	&	40763	\\
-%4(1)	&	22223	&	47771	&	&	135734	&	180489	\\
-%4(16)	&	1281	&	2483	&	&	9241	&	35024	\\
-%8(1)	&	13844	&	18933	&	&	77748	&	94563	\\
-%8(16)	&	608	&	1473	&	&	5820	&	23033	\\
-%12(1)	&	13844	&	18933	&	&	77748	&	94563	\\
-%12(16)	&	608	&	1473	&	&	5820	&	23033	\\
-		%\noalign{\smallskip}\hline
-	%\end{tabular}
-%\end{table}
 
@@
 	\center
-	\begin{tabular}{ccccccccc}
+	\begin{tabular}{cccccc}
 		\hline\noalign{\smallskip}
-		\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$}& & \multicolumn{2}{c}{$N=6$} \\
-		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \noalign{\smallskip}
-		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
+		\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
+		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
+		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
 		\noalign{\smallskip} \hline \noalign{\smallskip}
-1(1)	&	58320	&	84546	&	&	266943	&	287102 & & 1 & 2	\\
-1(16)	&	4297	&	6601	&	&	22655	&	56754	& &  & \\
-2(1)	&	34791	&	52126	&	&	188465	&	241369	& &  &\\
-2(16)	&	2029	&	3239	&	&	16689	&	40763	& &  &\\
-4(1)	&	22223	&	47771	&	&	135734	&	180489 & &  &	\\
-4(16)	&	1281	&	2483	&	&	9241	&	35024	& &  &\\
-8(1)	&	13844	&	18933	&	&	77748	&	94563	& &  &\\
-8(16)	&	608	&	1473	&	&	5820	&	23033	& &  &\\
-12(1)	&	13844	&	18933	&	&	77748	&	94563	& &  &\\
-12(16)	&	608	&	1473	&	&	5820	&	23033	& &  &\\
+1(1)	&	58320	&	84546	&	&	266943	&	287102	\\
+1(16)	&	4297	&	6601	&	&	22655	  &	56754	\\
+2(1)	&	34791	&	52126	&	&	188465	&	241369	\\
+2(16)	&	2029	&	3239	&	&	16689	  &	40763	\\
+4(1)	&	22223	&	47771	&	&	135734	&	180489	\\
+4(16)	&	1281	&	2483	&	&	9241	  &	35024	\\
+8(1)	&	13844	&	18933	&	&	77748	  &	94563	\\
+8(16)	&	608	  &	1473	&	&	5820	  &	23033	\\
+10(1)	&	13844	&	18933	&	&	77748	  &	94563	\\
+10(16)	&	608	&	1473	&	&	5820	  &	23033	\\
 		\noalign{\smallskip}\hline
@@
 Сложность моделируемых процессов и объектов может приводить к тому, что их характеристики не будут обладать свойством монотонности, что обуславливает недостаточность методов локального поиска для оценки наилучшего варианта.
-В таких задачах должны применяться методы глобального поиска, которые обеспечивают целенаправленность поиска за счет учета ограниченности изменения характеристик моделируемого объекта при ограниченных изменениях его параметров.
+В таких задачах должны применяться методы глобального поиска, которые обеспечивают целенаправленность поиска, например, за счет учета ограниченности изменения характеристик моделируемого объекта при ограниченных изменениях его параметров.
 Математическая формулировка этого факта может иметь форму Lipschitz condition, uniform H\"older condition и т.п.
@@
 \Russian
-Перспективным направлением дальнейшей работы будет являться применение множественных отображений для редукции размерности в задачах многокритериальной оптимизации.
+Перспективным направлением дальнейшей работы будет являться применение set of evolvents для редукции размерности в задачах многокритериальной оптимизации.
 В подобных задачах построение множества Парето требует решения целой серии скалярных задач.

### 2018-06-25 2018/Journal of Supercomputing/JoS_Paper.tex

@@
-%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%% file template.tex 
+%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%
 %
@@
 
-\usepackage[utf8x]{inputenc}
-\usepackage[english,russian]{babel}
+%\usepackage[utf8x]{inputenc}
+%\usepackage[english,russian]{babel}
 
@@
 \title{Parallel Computing for Constrained Global Optimization Problems
-\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}
+\thanks{The study was supported by the Russian Science Foundation, project No 16-11-
+10150.}
 }
@@
            I. Lebedev \at
-					Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
+					Lobachevsky State University of Nizhni Novgorod, Nizhni 
+Novgorod, Russia \\
 							\email{ilya.lebedev@itmm.unn.ru} 
@@
 \begin{abstract}
-This paper considers a parallel algorithm for solving multiextremal problems with non-convex constrains.
-Within the framework of the approach proposed here, the solution of the initial multidimensional problem is reduced to solving a set of linked one-dimensional problems in parallel.
-The parallelization scheme does not require a single control process since the solution of any one-dimensional problem is a solution of the initial multidimensional one.
-The non-convex constraints are accounted for by a novel index scheme without utilizing the ideas of penalty functions.
-The efficiency of the parallel algorithm was confirmed experimentally by solving numerically several hundred multidimensional multiextremal problems with non-convex constrains generated randomly.
-\keywords{Global optimization \and Constrained problems \and Non-convex constraints \and Dimensionality reduction \and Parallel algorithms}
+This paper considers a parallel algorithm for solving multiextremal problems with non-convex 
+constrains. Within the framework of the approach proposed here, the solution of the initial 
+multidimensional problem is reduced to solving a set of linked one-dimensional problems in 
+parallel. The parallelization scheme does not require a single control process since the solution of 
+any one-dimensional problem is a solution of the initial multidimensional one. The non-convex 
+constraints are accounted for by a novel index scheme without utilizing the ideas of penalty 
+functions. The efficiency of the parallel algorithm was confirmed experimentally by solving 
+numerically several hundred multidimensional multiextremal problems with non-convex 
+constrains generated randomly.
+\keywords{Global optimization \and Constrained problems \and Non-convex constraints \and 
+Dimensionality reduction \and Parallel algorithms}
 % \PACS{PACS code1 \and PACS code2 \and more}
@@
 
-The aim of this paper is to examine parallel methods for solving global optimization problems with non-convex constraints.
-The objective function and constraints are assumed to satisfy the Lipschitz condition with a priori unknown Lipschitz constants.
-The analytical form of the problem's functions may also be unknown, i.e.
-they can be set by an algorithm computing their values at various points within the search domain (the so-called ``black-box'' functions).
-It is also supposed that even a single computation of the function value can be time-consuming, since it requires performing numerical simulation in applied problems.
+The aim of this paper is to examine parallel methods for solving global optimization problems 
+with non-convex constraints. The objective function and constraints are assumed to satisfy the 
+Lipschitz condition with a priori unknown Lipschitz constants. The analytical form of the 
+problem's functions may also be unknown, i.e. they can be set by an algorithm computing their 
+values at various points within the search domain (the so-called ``black-box'' functions). It is 
+also supposed that even a single computation of the function value can be time-consuming, 
+since it requires performing numerical simulation in applied problems.
 
-The numerical algorithms for solving the problems of the above-mentioned class are noted for an exponential growth of the computational costs with increasing dimensionality (the so-called ``curse of dimensionality'').
-The use of supercomputers expands the field of application of the global optimization methods and, however, at the same time, it gives rise to the problem of efficient parallelization of the search process.
+The numerical algorithms for solving the problems of the above-mentioned class are noted for 
+an exponential growth of the computational costs with increasing dimensionality (the so-called 
+``curse of dimensionality''). The use of supercomputers expands the field of application of the 
+global optimization methods and, however, at the same time, it gives rise to the problem of 
+efficient parallelization of the search process.
 
-Parallel algorithms utilizing heuristic approaches and based in one way or another on the idea of random search are at present quite popular (see \cite{RefFerreiro,RefZhu,Korosec,Guerrero}).
-However, because of their probabilistic nature, the algorithms of such type guarantee the convergence to the global optimum solution in the probabilistic sense only, which makes them less attractive than deterministic parallel global optimization algorithms \cite{Jones2001,Paulavicius2011,Evtushenko2013,Strongin2000}.
-As a rule, efficient deterministic algorithms are intended to solve unconstrained optimization problems.
-The problems with constraints are solved using the penalty function method that has a number of disadvantages.
-In particular, this method cannot be applied to the problems with the \textit{partially defined functions} (i.e.
-when any constraint is violated, the values of all the other functions of the problem remain undefined).
+Parallel algorithms utilizing heuristic approaches and based in one way or another on the idea of 
+random search are at present quite popular (see \cite{RefFerreiro,RefZhu,Korosec,Guerrero}). 
+However, because of their probabilistic nature, the algorithms of such type guarantee the 
+convergence to the global optimum solution in the probabilistic sense only, which makes them 
+less attractive than deterministic parallel global optimization algorithms 
+\cite{Jones2001,Paulavicius2011,Evtushenko2013,Strongin2000}. As a rule, efficient 
+deterministic algorithms are intended to solve unconstrained optimization problems. The 
+problems with constraints are solved using the penalty function method that has a number of 
+disadvantages. In particular, this method cannot be applied to the problems with the 
+\textit{partially defined functions} (i.e. when any constraint is violated, the values of all the 
+other functions of the problem remain undefined).
 
-\Russian
-Данная ситуация является типичной для многих задач оптимального проектирования, когда при нарушении некоторых условий функционирования моделируемого объекта другие его характеристики оказываются неопределенными.
-Например, моделируемым объектом является некоторое электронное устройство, и минимизируется время его срабатывания.
-Однако время срабатывания является неопределенным, если при данном сочетании параметров устройство вообще не работает.
+This situation is typical for many optimal design problems, when at violation of some conditions 
+of functioning of the modeled object, other characteristics of this one appear to be undefined. 
+For example, the modeled object is an electronic device, and its response time is minimized. 
+However, the response time is undefined if at given combination of parameters the device is not 
+working at all.
 
-%Another example is the optimal control problem described through system of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}.
-It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e.
-every eigenvalue of $A$ has a strictly negative real part.
-Otherwise, the value of the criteria is indeterminate.
+%Another example is the optimal control problem described through system of ordinary 
+%differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}. It is only 
+%possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz 
+%matrix, i.e. every eigenvalue of $A$ has a strictly negative real part. Otherwise, the value of the 
+%criteria is indeterminate.
 
-In this study, we used a novel approach to the minimization of multiextremal functions with non-convex constraints developed in \cite{Strongin2000,Sergeyev2001,Barkalov2002}.
-The approach referred to as the index method of accounting for the constraints is based on accounting for each constraint of the problem separately and is not related to the use of the penalty functions.
-The solution of the initial multidimensional problem is reduced (by using the Peano-type space-filling curves) to solving an equivalent set of one-dimensional problems, which can be solved in parallel in such a way that each of these processes shares the information obtained by the other processes.
-The proposed scheme was implemented on the Lobachevsky supercomputer with the use of the distributed memory as well as the shared one.
+In this study, we used a novel approach to the minimization of multiextremal functions with 
+non-convex constraints developed in \cite{Strongin2000,Sergeyev2001,Barkalov2002}. The 
+approach referred to as the index method of accounting for the constraints is based on 
+accounting for each constraint of the problem separately and is not related to the use of the 
+penalty functions. The solution of the initial multidimensional problem is reduced (by using the 
+Peano-type space-filling curves) to solving an equivalent set of one-dimensional problems, 
+which can be solved in parallel in such a way that each of these processes shares the information 
+obtained by the other processes. The proposed scheme was implemented on the Lobachevsky 
+supercomputer with the use of the distributed memory as well as the shared one.
 
-The main part of the paper has the following structure.
-Section 2 states the constrained optimization problem, reviews the index method and an approach to reducing dimensionality by using Peano curves.
-Section 3 presents a parallel implementation of the index method using a set of space-filling curves.
-Section 4 presents the results of numerical experiments.
-Section 5 concludes the paper.
+The main part of the paper has the following structure. Section 2 states the constrained 
+optimization problem, reviews the index method and an approach to reducing dimensionality by 
+using Peano curves. Section 3 presents a parallel implementation of the index method using a set 
+of space-filling curves. Section 4 presents the results of numerical experiments. Section 5 
+concludes the paper.
 
@@
 \end{equation}
-The objective function $\varphi(y)$ (hereafter denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
+The objective function $\varphi(y)$ (hereafter denoted by $g_{m+1}(y)$) and the left-hand 
+sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
 \[
@@
 \]
-with a priori unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal.
-It is assumed that the functions $g_i(y)$ are defined and computable only at the points $y \in D$ satisfying the conditions
+with a priori unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal. It is 
+assumed that the functions $g_i(y)$ are defined and computable only at the points $y \in D$ 
+satisfying the conditions
 \begin{equation}\label{g_k}
@@
 
-By employing the continuous single-valued Peano curve $y(x)$ mapping the unit interval $[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}), it is possible to find the minimum in (\ref{problem}) by solving the one-dimensional problem
+By employing the continuous single-valued Peano curve $y(x)$ mapping the unit interval 
+$[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}), it is possible to find the 
+minimum in (\ref{problem}) by solving the one-dimensional problem
 \[
@@
 \]
-Algorithms for numerical construction of Peano curve approximation (\textit{evolvent}) are considered in \cite{Sergeyev2013}.
-Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined and computable in the subranges
+Algorithms for numerical construction of Peano curve approximation (\textit{evolvent}) are 
+considered in \cite{Sergeyev2013}. Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined 
+and computable in the subranges 
 \[
@@
 
-These conditions allows us to introduce a classification of the points $x \in [0,1]$ according to the number $\nu (x)$ of the constraints computed at this point.
-The \textit{index} $\nu(x)$ can also be defined by the conditions
+These conditions allows us to introduce a classification of the points $x \in [0,1]$ according to 
+the number $\nu (x)$ of the constraints computed at this point. The \textit{index} $\nu(x)$ can 
+also be defined by the conditions
 \begin{equation}\label{nu}
@@
 
-In the dimensionality reduction scheme considered here, a multidimensional problem with Lipschitzian functions is juxtaposed with a one-dimensional problem, where the corresponding functions satisfy uniform H{\"o}lder condition (see \cite{Sergeyev2013}), i.e.,
+In the dimensionality reduction scheme considered here, a multidimensional problem with 
+Lipschitzian functions is juxtaposed with a one-dimensional problem, where the corresponding 
+functions satisfy uniform H{\"o}lder condition (see \cite{Sergeyev2013}), i.e.,
 \[
-\left|g_i(y(x_1))-g_i (y(x_2))\right| \leq H_i \left|x_1-x_2 \right|^{1/N}, \; x_1,x_2\in [0,1], \; 1\leq i \leq m+1.
+\left|g_i(y(x_1))-g_i (y(x_2))\right| \leq H_i \left|x_1-x_2 \right|^{1/N}, \; x_1,x_2\in [0,1], \; 
+1\leq i \leq m+1.
 \]
-Here, $N$ is the dimensionality of the initial multidimensional problem and the coefficients $H_i$ are related to the Lipschitz constant $L_i$ of the initial problem as $H_i \leq 2L_i \sqrt{N+3}$.
+Here, $N$ is the dimensionality of the initial multidimensional problem and the coefficients 
+$H_i$ are related to the Lipschitz constant $L_i$ of the initial problem as $H_i \leq 2L_i 
+\sqrt{N+3}$.
 
-Thus, \textit{a trial} at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm will consist of the following sequence of operations:
+Thus, \textit{a trial} at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm 
+will consist of the following sequence of operations:
 \begin{itemize}
-	\item Determine the \textit{image} $y^k=y(x^k)$ in accordance with the mapping $y(x)$;
-	\item Compute the values $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from (\ref{nu}). 
+	\item Determine the \textit{image} $y^k=y(x^k)$ in accordance with the mapping 
+$y(x)$;
+	\item Compute the values $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from 
+(\ref{nu}). 
 \end{itemize}
-The dyad $ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} $ will be referred to as the \textit{trial outcome}.
+The dyad $ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} $ will be referred to as the \textit{trial 
+outcome}.
 
-The scheme of the serial index algorithm is as follows.
-The first trial is executed at an arbitrary internal point $x_1 \in (0,1)$.
-The selection of the point $x^{k+1}, \; k \geq 1,$ of any next trial is determined by the following steps.
+The scheme of the serial index algorithm is as follows. The first trial is executed at an arbitrary 
+internal point $x_1 \in (0,1)$. The selection of the point $x^{k+1}, \; k \geq 1,$ of any next trial 
+is determined by the following steps.
 
-Step 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the lower indices in the increasing order of the coordinate values, i.e.
+Step 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the lower indices in the 
+increasing order of the coordinate values, i.e.
 \[
@@
 \]
-and juxtapose to them the values $z_i=g_\nu(y(x_i)), \; \nu=\nu(x_i), \; 1 \leq i \leq k,$ computed at these points.
-The points $x_0=0$ and $x_{k+1}=1$ are introduced additionally and the values $z_0$ and $z_{k+1}$ are not defined.
+and juxtapose to them the values $z_i=g_\nu(y(x_i)), \; \nu=\nu(x_i), \; 1 \leq i \leq k,$ computed 
+at these points. The points $x_0=0$ and $x_{k+1}=1$ are introduced additionally and the 
+values $z_0$ and $z_{k+1}$ are not defined.
 
-Rule 2. For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} $R(i)$ using some formulae \cite{Barkalov2002}.
+Rule 2. For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} 
+$R(i)$ using some formulae \cite{Barkalov2002}.
 
@@
 
-Rule 4. Execute the next trial at the internal point of the interval $(x_{t-1},x_t)$, i.e. $x^{k+1} \in (x_{t-1},x_t)$.
+Rule 4. Execute the next trial at the internal point of the interval $(x_{t-1},x_t)$, i.e. $x^{k+1} 
+\in (x_{t-1},x_t)$.
 
-Step 5. Check termination condition $\left|x_t-x_{t-1}\right|^{1/N}\leq \epsilon$, where $t$ is the number of interval with the maximum characteristic and $\epsilon > 0$ is the predefined accuracy.
+Step 5. Check termination condition $\left|x_t-x_{t-1}\right|^{1/N}\leq \epsilon$, where $t$ is 
+the number of interval with the maximum characteristic and $\epsilon > 0$ is the predefined 
+accuracy.
 
-A detailed description of this algorithm and the corresponding theory of convergence are presented in \cite{Strongin2000,Sergeyev2001,Barkalov2002}.
+A detailed description of this algorithm and the corresponding theory of convergence are 
+presented in \cite{Strongin2000,Sergeyev2001,Barkalov2002}.
 
-\Russian
-Сравнение двух разных способа учета ограничений - классического метода штрафных функций и индексной схемы - приведено в статье \cite{Barkalov2017}.
-В ней экспериментально подтрвеждено превосходство индексного метода над методом штрафных функций как по числу итераций, так и по числу вычсислений ограничений и целевой функции.
-Для решения задачи с заданной точностью индексному методу требовалось в 3-5 раз меньше итераций (в зависимости от размерности задачи и доли допусимой области).
-Одновременно с этим значение целевой функции вычислялось в среднем лишь на каждой пятой итерации индексного метода, в отличие от метода штрафных функций, в котором значения целевой функции вычисляются на каждой итерации.
+The comparison of two different methods of accounting for the constraints -- the classical 
+penalty functions method and the index scheme -- is presented in \cite{Barkalov2017}. In 
+this paper, the advantage of the index method over the penalty functions one in the number of 
+iterations as well as in the number of computations of the constraints and of the objective 
+function has been confirmed experimentally. To solve a problem with given accuracy, the index 
+method required 3-5 times less iterations (subject to the problem dimensionality and the feasible 
+domain fraction). At the same time, the objective function value was computed in average at 
+every fifth iteration of the index method only, in difference from the penalty functions method, 
+in which the objective function values were computed at every iteration.
 
@@
 
-The reduction of the multidimensional problems to the one-dimensional ones using evolvents has such important properties as the continuity and preservation of boundedness of function's divided differences.
-However, a partial loss of information on the nearness of the points in the multidimensional space occurs in this case since a point $x \in [0,1]$ has only the left and the right neighbors while the corresponding point $y(x) \in R^N$ has the neighbors in $2N$ directions.
-As a result, when using the mappings like a Peano curve, the close images $y', y'' \in D$ can correspond to the remote preimages $x', x'' \in [0,1]$.
-This property results in excess computations since several limit points $x', x''$ of the trial sequence generated by the index method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional space.
+The reduction of the multidimensional problems to the one-dimensional ones using evolvents 
+has such important properties as the continuity and preservation of boundedness of function's 
+divided differences. However, a partial loss of information on the nearness of the points in the 
+multidimensional space occurs in this case since a point $x \in [0,1]$ has only the left and the 
+right neighbors while the corresponding point $y(x) \in R^N$ has the neighbors in $2N$ 
+directions. As a result, when using the mappings like a Peano curve, the close images $y', y'' \in 
+D$ can correspond to the remote preimages $x', x'' \in [0,1]$. This property results in excess 
+computations since several limit points $x', x''$ of the trial sequence generated by the index 
+method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional 
+space.
 
-One of the possible ways to overcome this disadvantage consists in using the multiple mapping $Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of the single evolvent $y(x)$.
-To construct the set $Y^S(x)$, different approaches can be used.
-%For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^s(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$.
-The set of Peano curves thus constructed allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$, for any close multidimensional images $y', y''$ from $D$, which differ only in one coordinate.
+One of the possible ways to overcome this disadvantage consists in using the multiple mapping 
+$Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of the single evolvent $y(x)$. To construct the set 
+$Y^S(x)$, different approaches can be used. 
+%For example, in \cite{Strongin2000} a scheme was implemented, according to which each 
+%evolvent $y^s(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent 
+%$y(x)$ along the main diagonal of the hypercube $D$. The set of Peano curves thus constructed 
+%allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 
+%1\leq s \leq S$, for any close multidimensional images $y', y''$ from $D$, which differ only in 
+%one coordinate.
 
-%Проверить
-For example, a scheme of building of the multiple mapping was proposed  in \cite{6_Gergel2009}.
-The building of a Peano curves set by rotation of the evolvents around the coordinate origin is a distinctive feature of this scheme.
-In the initial non-rotated evolvent for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other.
-In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer.
-The maximum number of various rotations of the evolvents mapping the $N$-dimensional hypercube onto a one-dimensional interval is $2^N$.
-The employment of all possible rotations might appear to be redundant.
-In this case, we can select only a part of all rotations.
-As a possible approach, we propose to generate the new evolvents by the rotation of the initial evolvent on the angle of $\pm\pi/2$ in each of the coordinate planes.
-Taking into account the initial evolvent the total number of such evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals equals to $N(N-1)+1$.
-This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.
-\Russian
-Именно такой подход к построению set of evolvents был использован в данной статье при проведении вычислительных экспериментов.
+For example, a scheme of building of the multiple mapping was proposed in \cite{6_Gergel2009}. The building of a set of Peano curves by rotation of the evolvents 
+around the coordinate origin is a distinctive feature of this scheme. In the initial non-rotated 
+evolvent for close points $y', y''$ in the multidimensional space their preimages $x', x''$ in the 
+interval $[0,1]$ can be far away from each other. In the rotated scheme there exists a mapping 
+$y^i(x)$ according to which the preimages $x', x''$ will be located closer to each other. The 
+maximum number of various rotations of the evolvents mapping the $N$-dimensional 
+hypercube onto a one-dimensional interval is $2^N$. The employment of all possible rotations 
+might appear to be redundant. In this case, we can select only a part of all rotations. As a 
+possible approach, we propose to generate the new evolvents by the rotation of the initial 
+evolvent on the angle of $\pm\pi/2$ in each coordinate plane. Taking into account the initial 
+evolvent, the total number of such evolvents for mapping the $N$-dimensional domain onto the 
+corresponding one-dimensional intervals equals to $N(N-1)+1$. This method for building a set 
+of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.
+Just such an approach to building the set of evolvents was employed in the present paper when 
+carrying out the numerical experiments.
 
-With the use of multiple mapping, it is possible to solve the initial problem (\ref{problem}) by solving in parallel the problems
+With the use of multiple mapping, it is possible to solve the initial problem (\ref{problem}) by 
+solving in parallel the problems
 \[
@@
 \]
-on a set of intervals $[0,1]$ by the index method.
-Each one-dimensional problem is solved on a separate processor/node.
-The trial outcome at the point $x^k$ obtained for the problem being solved by particular processor are interpreted as the outcomes of the trials in the remaining problems (at the corresponding points $x^{k_1},...,x^{k_S}$).
-In this approach, a trial at the point $x^k \in [0,1]$ executed in the framework of the $s$-th problem, consists of the following sequence of operations.
+on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a 
+separate processor/node. The trial outcome at the point $x^k$ obtained for the problem being 
+solved by particular processor are interpreted as the outcomes of the trials in the remaining 
+problems (at the corresponding points $x^{k_1},...,x^{k_S}$). In this approach, a trial at the 
+point $x^k \in [0,1]$ executed in the framework of the $s$-th problem, consists of the following 
+sequence of operations.
 
@@
 
-	\item Inform the rest of processors about the start of the trial execution at the point $y^k$ (\textit{the blocking} of the point $y^k$).
+	\item Inform the rest of processors about the start of the trial execution at the point 
+$y^k$ (\textit{the blocking} of the point $y^k$).
 
-	\item Compute the values $g_1(y^k),...,g_\nu(y^k)$, where the index $\nu \leq m$ is determined by the conditions
+	\item Compute the values $g_1(y^k),...,g_\nu(y^k)$, where the index $\nu \leq m$ is 
+determined by the conditions
 	\[
@@
 	\]
-The trial terminates at the point as soon as the first violation of any constraint occurs.
-In the case when $y^k$ is a feasible one, the trial includes the computation of the values of all problem functions.
-In this situation, the index is set to $\nu=m+1$.
-The triplet
+The trial terminates at the point as soon as the first violation of any constraint occurs. In the case 
+when $y^k$ is a feasible one, the trial includes the computation of the values of all problem 
+functions. In this situation, the index is set to $\nu=m+1$. The triplet
 \[
@@
 
-\item Determine the preimages $x^{k_s} \in [0,1]$, $1 \leq s \leq S$, of the point $y^k$ and interpret the trial executed at the point $y^k \in D$ as the execution of the trials at the $S$ points  $x^{k_1},...,x^{k_S}$ with the same outcomes
+\item Determine the preimages $x^{k_s} \in [0,1]$, $1 \leq s \leq S$, of the point $y^k$ and 
+interpret the trial executed at the point $y^k \in D$ as the execution of the trials at the $S$ 
+points  $x^{k_1},...,x^{k_S}$ with the same outcomes
 \[
@@
 
-The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the sequential algorithm (with the exception the method of the trial execution).
-To organize the interactions between the processors, the queues are created on each processor, where the processors store the information on the executed iterations in the form of tuples: the processor number $s$, the trial point $x^{k_s}$, the index $\nu(x^{k_s})$, and the value $g_\nu(y^s (x^{k_s} ) )$.
-Moreover, the index of the blocked point is assumed to be equal to $-1$; the function value at this point is undefined.
+The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the 
+sequential algorithm (with the exception the method of the trial execution). To organize the 
+interactions between the processors, the queues are created on each processor, where the 
+processors store the information on the executed iterations in the form of tuples: the processor 
+number $s$, the trial point $x^{k_s}$, the index $\nu(x^{k_s})$, and the value $g_\nu(y^s
+(x^{k_s} ) )$. Moreover, the index of the blocked point is assumed to be equal to $-1$; the 
+function value at this point is undefined.
 
-The proposed parallelization scheme was implemented with the use of MPI and OpenMP technologies.
-A separate MPI-process is created for each of $S$ one-dimensional problems being solved, one process per one processor employed.
-Each process uses $p$ threads for solving a one-dimensional problem, one thread per an accessible core.
+The proposed parallelization scheme was implemented with the use of MPI and OpenMP 
+technologies. A separate MPI-process is created for each of $S$ one-dimensional problems 
+being solved, one process per one processor employed. Each process uses $p$ threads for 
+solving a one-dimensional problem, one thread per an accessible core.
 
@@
 
-A well-known approach to investigating and comparing the multiextremal optimization algorithms is based on testing these methods by solving a set of problems, chosen randomly from some specially designed class.
+A well-known approach to investigating and comparing the multiextremal optimization 
+algorithms is based on testing these methods by solving a set of problems, chosen randomly 
+from some specially designed class.
 
-In this study, we will use a GKLS generator of the functions of arbitrary dimensionality proposed in \cite{Gaviano2003} to produce constrained problems.
-The scheme for generating constrained global optimization problems is proposed in \cite{Gergel2017}.
-\Russian
-Тестовые задачи с ограничениями, полученные с помощью данной схемы, обладают важным свойством, присущим прикладным задачам оптимизации: условный глобальный минимум находится на границе допустимой области, в то время как глобальный минимум целевой функции без ограничений находится вне допустимой области.
-При этом можно задавать объем допустимой области и  число ограничений, активных в точке минимума, тем самым усложняя или упрощая постановку задачи.
+In this study, we will use a GKLS generator of the functions of arbitrary dimensionality 
+proposed in \cite{Gaviano2003} to produce constrained problems. The scheme for generating 
+constrained global optimization problems is proposed in \cite{Gergel2017}. 
+The test problems with constraints obtained using this scheme have an important property 
+inherent to the applied optimization problems: the constrained global minimum is located at the 
+feasible domain boundary whereas the global minimum of the objective function without 
+constraints is located outside the feasible domain. At that, one can specify the feasible domain 
+volume and the number of constraints active at the minimum point thus complicating or 
+simplifying the problem statement.
 
-In the present study, an experimental investigation is carried out of the speedup, which is obtained by the use of the index method in combination with the two-level parallelization scheme from \cite{Sidorov2015}.
+In the present study, an experimental investigation is carried out of the speedup, which is 
+obtained by the use of the index method in combination with the two-level parallelization 
+scheme from \cite{Sidorov2015}.
 
-The experiments have been carried out by solving a series of $100$ problems with two constraints and one objective function from the \textit{Simple} and \textit{Hard} GKLS classes with the dimensionalities $N=4$, $N=5$.
-The number of the used cluster nodes $S$ and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at each node have been varied.
-The problem was considered to be solved, if the algorithm generated a trial point $y^k$ in the $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^*\right\|\leq \delta = 0.03\left\|b-a\right\|$.
-For the purpose of simulation of the computational complexity inherent to applied optimization problems, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (a series summation of $100$ thousand elements).
+The experiments have been carried out by solving a series of $100$ problems with two 
+constraints and one objective function from the \textit{Simple} and \textit{Hard} GKLS 
+classes with the dimensionalities $N=4$, $N=5$. The number of the used cluster nodes $S$ 
+and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at 
+each node have been varied. The problem was considered to be solved, if the algorithm 
+generated a trial point $y^k$ in the $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-
+y^*\right\|\leq \delta = 0.03\left\|b-a\right\|$. For the purpose of simulation of the computational 
+complexity inherent to applied optimization problems, calculation of the problem functions in all 
+performed experiments was made more complex by additional calculations without changing 
+the type of function and arrangement of its minima (a series summation of $100$ thousand 
+elements).
 
-Computational experiments were carried out on Lobachevsky supercomputer.
-The node of supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.
-The CPU had 8 cores, i.e.
-each node had a total of 16 cores.
-The average time and the number of iterations, which were required to solve the problems of the series at various parallelization parameters are given in Tables 1 and 2.
-Here, \textit{Node(core)} are the numbers of employed  nodes and cores per node, respectively.
+Computational experiments were carried out on Lobachevsky supercomputer. The node of 
+supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The 
+CPU had 8 cores, i.e. each node had a total of 16 cores. The average time and the number of 
+iterations, which were required to solve the problems of the series at various parallelization 
+parameters are given in Tables 1 and 2. Here, \textit{Node(core)} are the numbers of employed  
+nodes and cores per node, respectively.
 
@@
 		\hline\noalign{\smallskip}
-		\multirow{2}{*}{\textit{Node(core)}} & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
+		\multirow{2}{*}{\textit{Node(core)}} & \multicolumn{2}{c}{ $N=4$ } & & 
+\multicolumn{2}{c}{$N=5$} \\
 		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
@@
 		\hline\noalign{\smallskip}
-		\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
+		\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & 
+& \multicolumn{2}{c}{$N=5$} \\
 		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
@@
 
-The results demonstrate the speedup achieved when using the shared memory at a node as well as the distributed memory.
-In this case, the highest time speedup was 10 (when using 64 cores on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster nodes).
-The difference in the speedups in time and in the number of iterations can be explained by the effect of the overheads of the data transmission between the processes.
-Note that when solving applied optimization problems, the computation of the problem function values even in one point is a computationally costly operation.
-In this case, the data transfer overheads will not affect significantly the total computational costs, and the difference between the time speedup and the iteration speedup will not be so great.
+The results demonstrate the speedup achieved when using the shared memory at a node as well 
+as the distributed memory. In this case, the highest time speedup was 10 (when using 64 cores 
+on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster 
+nodes). The difference in the speedups in time and in the number of iterations can be explained 
+by the effect of the overheads of the data transmission between the processes. Note that when 
+solving applied optimization problems, the computation of the problem function values even in 
+one point is a computationally costly operation. In this case, the data transfer overheads will not 
+affect significantly the total computational costs, and the difference between the time speedup 
+and the iteration speedup will not be so great.
 
@@
 
-\Russian
-В заключение отметим, что подавляющее большинство оптимизационных алгоритмов (за исключением методов, основанных на идеях случайного поиска) реализуют стратегию целенаправленного выбора.
-Указанная стратегия состоит в том, чтобы на основе анализа небольшой части вариантов исключить из дальнейшего рассмотрения многие заведомо не перспективные случаи и сосредоточить дальнейший поиск в подмножестве, содержащем лучший вариант.
+In conclusion, note that the majority of the optimization algorithms (except the methods based 
+on the random search ideas) implement the purposeful search strategy. This strategy consists in 
+excluding many non-prospective cases from further consideration on the base of 
+analysis of a small part of variants and concentrating further search in the subset containing the 
+best variant.
 
-Сложность моделируемых процессов и объектов может приводить к тому, что их характеристики не будут обладать свойством монотонности, что обуславливает недостаточность методов локального поиска для оценки наилучшего варианта.
-В таких задачах должны применяться методы глобального поиска, которые обеспечивают целенаправленность поиска, например, за счет учета ограниченности изменения характеристик моделируемого объекта при ограниченных изменениях его параметров.
-Математическая формулировка этого факта может иметь форму Lipschitz condition, uniform H\"older condition и т.п.
-Именно к таким методам, способным решать сложные прикладные задачи оптимизации, относится рассмотренный в данной статье параллельный индексный алгоритм.
+The complexity of the modeled processes and objects can result in the non-monotonous
+characteristics of these ones that causes the inadequacy of local search 
+methods for evaluating the optimal variant. In such problems, the global search methods should 
+be applied, which provide the purposeful search, for example, by means of accounting for the 
+limited variation of the characteristics of the modeled object at limited variations of its 
+parameters. The mathematical formulation of this fact can take the forms of the Lipschitz 
+condition, uniform H\"older condition, etc. The parallel index algorithm considered in the 
+present paper belongs just to such methods capable to solve the complex applied optimization 
+problems.
 
-The parallel index method for solving constrained global optimization problems %considered in this paper 
+The parallel index algorithm %for solving constrained global optimization problems %considered in this paper 
 offers the following possibilities:
 \begin{itemize}
-	\item to solve the initial problem directly, without the use of the penalty functions (thus eliminating the need to select the penalty coefficient and to solve a series of unconstrained problems with different penalty coefficients);
-	\item to solve the problems where the values of the problem function are not defined everywhere (for example, the objective function values are undefined outside the feasible domain);
-	\item to use the two-level parallelization scheme with the shared and distributed memory proposed earlier for the unconstrained optimization methods.
+	\item to solve the initial problem directly, without the use of the penalty functions (thus 
+eliminating the need to select the penalty coefficient and to solve a series of unconstrained 
+problems with different penalty coefficients);
+	\item to solve the problems where the values of the problem function are not defined 
+everywhere (for example, the objective function values are undefined outside the feasible 
+domain);
+	\item to use the two-level parallelization scheme with the shared and distributed memory 
+proposed earlier for the unconstrained optimization methods.
 \end{itemize}
 
-The parallel algorithm has demonstrated a speedup with respect to the number of processors/cores employed.
-This was confirmed by the results of solving numerically of several hundred test problems using 128 cores of the Lobachevsky supercomputer.
+The parallel algorithm has demonstrated a speedup with respect to the number of 
+processors/cores employed. This was confirmed by the results of solving numerically of several 
+hundred test problems using 160 cores of the Lobachevsky supercomputer.
 
-\Russian
-Перспективным направлением дальнейшей работы будет являться применение set of evolvents для редукции размерности в задачах многокритериальной оптимизации.
-В подобных задачах построение множества Парето требует решения целой серии скалярных задач.
-При их параллельном решении можно ожидать более высокого ускорения, чем при решении одной скалярной задачи.
+The application of the set of evolvents for the dimensionality reduction in the multicriterial 
+optimization problems would be a promising direction of further work. In such problems, the 
+construction of Pareto set requires solving even a series of scalar problems. When solving these 
+ones in parallel, one can expect a higher speedup than when solving a single scalar problem.
 
@@
 \bibitem{RefFerreiro}
-Ferreiro, A.M., Garcia, J.A., Lopez-Salas, J.G., Vazquez, C.: An efficient implementation of parallel simulated annealing algorithm in GPUs. J. Glob. Optim. 57(3), 863--890 (2013)
+Ferreiro, A.M., Garcia, J.A., Lopez-Salas, J.G., Vazquez, C.: An efficient implementation of 
+parallel simulated annealing algorithm in GPUs. J. Glob. Optim. 57(3), 863--890 (2013)
 
 \bibitem{RefZhu}
-Zhu, W.: Massively parallel differential evolution--pattern search optimization with graphics hardware acceleration: an investigation on bound constrained optimization problems.
-J.
-Glob.
-Optim.
-50(3), 417--437 (2011).
+Zhu, W.: Massively parallel differential evolution--pattern search optimization with graphics 
+hardware acceleration: an investigation on bound constrained optimization problems. J. Glob. 
+Optim. 50(3), 417--437 (2011).
 
 \bibitem{Korosec}
-Koro\v sec, P., Vajter\v sic, M.,  \v Silc, J., Kutil, R.: Multi-core implementation of the differential ant-stigmergy algorithm for numerical optimization. J. Supercomput. 63(3), 757--772 (2013)
+Koro\v sec, P., Vajter\v sic, M.,  \v Silc, J., Kutil, R.: Multi-core implementation of the 
+differential ant-stigmergy algorithm for numerical optimization. J. Supercomput. 63(3), 757--
+772 (2013)
 
 \bibitem{Guerrero}
-Guerrero, G.D., Cecilia, J.M., Llanes, A. et al.: Comparative evaluation of platforms for parallel Ant Colony Optimization. J. Supercomput. 69(1) 318--329 (2014)
+Guerrero, G.D., Cecilia, J.M., Llanes, A. et al.: Comparative evaluation of platforms for parallel 
+Ant Colony Optimization. J. Supercomput. 69(1) 318--329 (2014)
 
 \bibitem{Jones2001}
-Jones, D.R.: The direct global optimization algorithm. In: Floudas, C.A., Pardalos, P.M. (eds.) The Encyclopedia of Optimization, Second Edition. pp. 725--735. Springer (2009)
+Jones, D.R.: The direct global optimization algorithm. In: Floudas, C.A., Pardalos, P.M. (eds.) 
+The Encyclopedia of Optimization, Second Edition. pp. 725--735. Springer (2009)
 
 \bibitem{Paulavicius2011}
-Paulavi\v cius, R., \v Zilinskas, J., Grothey, A.: Parallel branch and bound for global optimization with combination of Lipschitz bounds. Optim. Methods Softw. 26(3), 487--498 (2011)
+Paulavi\v cius, R., \v Zilinskas, J., Grothey, A.: Parallel branch and bound for global 
+optimization with combination of Lipschitz bounds. Optim. Methods Softw. 26(3), 487--498 
+(2011)
 
 \bibitem{Evtushenko2013}
-Evtushenko, Y.G., Posypkin, M.A.: A deterministic approach to global box-constrained optimization. Optim. Lett. 7(4), 819--829 (2013)
+Evtushenko, Y.G., Posypkin, M.A.: A deterministic approach to global box-constrained 
+optimization. Optim. Lett. 7(4), 819--829 (2013)
 
 \bibitem{Strongin2000}
-Strongin, R.G., Sergeyev, Ya.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000)
+Strongin, R.G., Sergeyev, Ya.D.: Global Optimization with Non-Convex Constraints. 
+Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000)
 
-%\bibitem{Balandin2011} Balandin, D.V., Kogan, M.M.: Optimal linear-quadratic control: From matrix equations to linear matrix inequalities. Automation and Remote Control. 72(11), 2276--2284 (2011)
+%\bibitem{Balandin2011} Balandin, D.V., Kogan, M.M.: Optimal linear-quadratic control: 
+From matrix equations to linear matrix inequalities. Automation and Remote Control. 72(11), 
+2276--2284 (2011)
 
-%\bibitem{Balandin2017}Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 (2017)
+%\bibitem{Balandin2017}Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-
+control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 
+(2017)
 
 \bibitem{Sergeyev2001}
-Sergeyev, Y.D., Famularo D., Pugliese P. Index Branch-and-Bound Algorithm for Lipschitz univariate global optimization with multiextremal constraints. J. Glob. Optim. 21(3), 317--341 (2001)
+Sergeyev, Y.D., Famularo D., Pugliese P. Index Branch-and-Bound Algorithm for Lipschitz 
+univariate global optimization with multiextremal constraints. J. Glob. Optim. 21(3), 317--341 
+(2001)
 
 \bibitem{Barkalov2002}
-Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)
+Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of 
+checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)
 
 \bibitem{Sergeyev2013}
-Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting Space-Filling Curves. Springer (2013)
+Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting 
+Space-Filling Curves. Springer (2013)
 
 \bibitem{Gaviano2003}
-Gaviano, M., Lera, D., Kvasov, D.E., Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
-ACM Trans.
-Math.
+Gaviano, M., Lera, D., Kvasov, D.E., Sergeyev, Ya.D.: Software for generation of classes of 
+test functions with known local and global minima for global optimization. ACM Trans. Math. 
 Software 29, 469--480 (2003)
@@
 \bibitem{Gergel2017}
-Gergel, V.: An approach for generating test problems of constrained global optimization. LNCS. 10556, 314--319 (2017)
+Gergel, V.: An approach for generating test problems of constrained global optimization. 
+LNCS. 10556, 314--319 (2017)
 
-%\bibitem{Barkalov2015} Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi coprocessor for solving global optimization problems. LNCS. 9251, 307--318 (2015)
+%\bibitem{Barkalov2015} Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi coprocessor 
+for solving global optimization problems. LNCS. 9251, 307--318 (2015)
 
 \bibitem{Sidorov2015}
-Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of computationally intensive multiextremal optimization problems. LNCS. 9251, 505--515 (2015)
+Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of 
+computationally intensive multiextremal optimization problems. LNCS. 9251, 505--515 (2015)
 
 \bibitem{Barkalov2017}
-Barkalov, K., Lebeled, I.: Comparing Two Approaches for Solving Constrained Global Optimization Problems. LNCS. 10556, 301--306 (2017)
+Barkalov, K., Lebeled, I.: Comparing Two Approaches for Solving Constrained Global 
+Optimization Problems. LNCS. 10556, 301--306 (2017)
 
 \bibitem{6_Gergel2009}
-Strongin, R.G., Gergel, V.P., Barkalov, K.A.: Parallel methods for global optimization problem solving. Journal of instrument engineering. 52, 25--33 (2009) (In Russian)
+Strongin, R.G., Gergel, V.P., Barkalov, K.A.: Parallel methods for global optimization problem 
+solving. Journal of instrument engineering. 52, 25--33 (2009) (In Russian)
 

### 2018-06-26 2018/YSC/slides/slides.tex
YSC: fix year at the first slide
YSC: update slides

@@
 
+\captionsetup[subfigure]{labelformat=empty}
+\captionsetup[figure]{labelformat=empty}
+
 \setromanfont{CMU Serif}
@@
 
-\graphicspath{ {../paper/images/}{img/} }
+\graphicspath{ {../pictures/}{img/} }
 %set pages numeration
@@
 \begin{frame}
-  \begin{center}
   \frametitle{Problem statement}
-
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
       \begin{displaymath}
-    \min\{f(y): y\in D\}, D=\{y\in \mathbb{R}^n: a_i \leqslant y_i \leqslant b_i, 1\leqslant i \leqslant n \},
+        \begin{array}{cr}\\
+          \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
+          D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
+        \end{array}
       \end{displaymath}
-where \(f(y)\) is a vector-function.
-
-\enspace
-Solution of the problem is a set of non-dominated points (Slater set):
+      \(\varphi(y)\) is multiextremal objective function, which satisfies the Lipschitz condition:
       \begin{displaymath}
-    S(D) = \{y\in D: \nexists z\in D, f_i(z)<f_i(y),1\leqslant i \leqslant m\}
+        |\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
       \end{displaymath}
-Assume objectives to satisfy Lipschitz condition in \(D\):
-  \begin{displaymath}
-    |f_i(y_1)-f_i(y_2)|\leqslant L_i\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L_i<\infty,1\leqslant i\leqslant m
-  \end{displaymath}
-
-\end{center}
+      where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes \(l_2\) norm in \(\mathbb{R}^N\)
+      space.
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \centerline{\includegraphics[width=0.9\textwidth]{img/gkls.png}}
+    \end{column}
+  \end{columns}
 \end{frame}
@@
   \frametitle{Dimension reduction}
-  Peano-type curve \(y(x)\) allows to reduce dimension of the original multi-objective problem:
+  Peano-type curve \(y(x)\) allows to reduce the dimension of the original problem:
   \begin{gather}
@@
 
-  %\begin{figure}[ht]
-  %  \subfloat{{\includegraphics[width=.35\textwidth]{img/peano3d.png} }}
-  %  \subfloat{{\includegraphics[width=.35\textwidth]{img/peano2d.png} }}
-  %\end{figure}
+  \begin{figure}[ht]
+    \subfloat{{\includegraphics[width=.35\textwidth]{peano3d.png} }}
+    \subfloat{{\includegraphics[width=.35\textwidth]{peano2d.png} }}
+  \end{figure}
+\end{frame}
+
+\begin{frame}
+  \frametitle{}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Smooth evolvent}
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \centerline{\includegraphics[width=0.9\textwidth]{smooth.pdf}}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Shifted evolvent}
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \centerline{\includegraphics[width=0.9\textwidth]{shifted.pdf}}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Rotated evolvent}
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \centerline{\includegraphics[width=0.9\textwidth]{rotated.pdf}}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Optimization method}
+  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
+  \begin{enumerate}
+    \setlength{\itemindent}{.1in}
+    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
+    \item[Step 2.] Compute the evolvent \(y(x)\) and the function \(\varphi(y(x))\).
+    \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
+    \item[Step 4.] Choose an interval \((x_{t-1}, x_t)\) with the greatest characteristic and
+    compute objective \(f(y(x))\) in the point chosen using the decision rule \(d\):
+    \begin{displaymath}
+      x^{k+1}=d(t)\in (x_{t-1}, x_t)
+    \end{displaymath}
+    \item[Step 5.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
+  \end{enumerate}
+  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Test problems}
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+      Generator GKLS was employed to construct the sets of test problems:
+      \begin{displaymath}
+        \begin{matrix}
+          f(x)=
+          \left\{
+          \begin{matrix}
+          C_i(x), x \in S_i, i\in 2,\dots ,m \\
+          \Vert x-T \Vert^2 + t, x\not\in S_2,\dots,S_m
+          \end{matrix} \right.
+        \end{matrix}
+      \end{displaymath}
+
+      The generator allows to adjust:
+      \begin{itemize}
+        \item the number of local minimas;
+        \item the size of the global minima attraction region;
+        \item the space dimension.
+      \end{itemize}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \centerline{\includegraphics[width=0.9\textwidth]{gkls_color.png}}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results of comparison}
+  \begin{figure}[ht]
+    \hspace*{-0.9cm}
+    \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_opt_pt_op.pdf} }}
+    \subfloat[\(r=5.0\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_same_r_opt_pt_op.pdf} }} \hspace*{4cm}
+    \caption{Operating characteristics on GKLS 2d Simple class}
+  \end{figure}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results of comparison}
+  \begin{figure}[ht]
+    \hspace*{-0.9cm}
+    \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_opt_pt_op.pdf} }}
+    \subfloat[\(r=4.5\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_same_r_opt_pt_op.pdf} }}
+    \caption{Operating characteristics on GKLS 3d Simple class}
+  \end{figure}
 \end{frame}
@@
     \begin{itemize}
+      \item
     \end{itemize}

### 2018-06-27 2018/YSC/slides/slides.tex
YSC: update slides

@@
 \usepackage{subfig}
+\usepackage{tikz}
 
@@
 \begin{frame}
+  \begin{center}
+
   \frametitle{Dimension reduction}
@@
   \end{gather}
+  \(y(x)\) is non-smooth function which continuously maps the segment \([0,1]\) to the hypercube \(D\).
+  \begin{figure}[ht]
+    \vspace*{-0.5cm}
+    \subfloat{{\includegraphics[width=.35\textwidth]{peano3d.png} }}
+    \subfloat{{\includegraphics[width=.35\textwidth]{peano2d.png} }}
+  \end{figure}
+\end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Properties of the reduced problem}
+  After applying the Peano-type evolvent \(\varphi(y(x))\) satisfies the uniform H{\"o}lder condition:
+  \begin{displaymath}
+    |\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
+  \end{displaymath}
+  \(\varphi(y(x))\) is non-smooth and has multiple local extremums even if \(\varphi(y)\) is unimodal.
+  The latter problem is caused by loss of the information about \(N\)-d neighborhood after the transformation to the \(1\)-d space.
 
   \begin{figure}[ht]
-    \subfloat{{\includegraphics[width=.35\textwidth]{peano3d.png} }}
-    \subfloat{{\includegraphics[width=.35\textwidth]{peano2d.png} }}
+    \begin{center}
+
+    \vspace*{-0.5cm}
+    \subfloat{\raisebox{.02\textheight}{\includegraphics[width=.3\textwidth]{parabaloid.png}}}
+    \subfloat{\raisebox{.2\textheight}{\includegraphics[width=.05\textwidth]{arrow.jpg}}}
+    \subfloat{{\includegraphics[width=.45\textwidth]{map_paraboloid.pdf}}}
+  \end{center}
   \end{figure}
@@
 \begin{frame}
-  \frametitle{}
-
+  \frametitle{Smooth evolvent and non-univalent evolvent}
+      %\centerline{\includegraphics[width=0.9\textwidth]{smooth.pdf}}
+      \begin{figure}[ht]
+        %\vspace*{-0.5cm}
+        \subfloat{\raisebox{.02\textheight}{\includegraphics[width=.35\textwidth]{smooth.pdf} }}
+        \subfloat{{\includegraphics[width=.25\textwidth]{noninjective.pdf} }}
+      \end{figure}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Smooth evolvent}
-  \begin{columns}
-    \begin{column}{0.5\textwidth}
-
-    \end{column}
-    \begin{column}{0.5\textwidth}
-      \centerline{\includegraphics[width=0.9\textwidth]{smooth.pdf}}
-    \end{column}
-  \end{columns}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Shifted evolvent}
-  \begin{columns}
-    \begin{column}{0.5\textwidth}
-
-    \end{column}
-    \begin{column}{0.5\textwidth}
-      \centerline{\includegraphics[width=0.9\textwidth]{shifted.pdf}}
-    \end{column}
-  \end{columns}
+  \frametitle{Shifted and rotated evolvents}
+  \begin{figure}[ht]
+    %\vspace*{-0.5cm}
+    \subfloat{\raisebox{.0\textheight}{\includegraphics[width=.35\textwidth]{shifted.pdf} }}
+    \subfloat{{\includegraphics[width=.35\textwidth]{rotated.pdf} }}
+  \end{figure}
+%  \begin{columns}
+%    \begin{column}{0.5\textwidth}
+%
+%    \end{column}
+%    \begin{column}{0.5\textwidth}
+%      \centerline{\includegraphics[width=0.9\textwidth]{shifted.pdf}}
+%    \end{column}
+%  \end{columns}
 \end{frame}
@@
     \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_opt_pt_op.pdf} }}
-    \subfloat[\(r=4.5\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_same_r_opt_pt_op.pdf} }}
+    \subfloat[\(r=4.5\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_same_r_opt_pt_op.pdf} }}
     \caption{Operating characteristics on GKLS 3d Simple class}

### 2018-07-10 2018/YSC/ysc_paper.tex
YSC: update to the new procedia template

@@
-% easychair.tex,v 3.1 2011/12/30
-%
-% Select appropriate paper format in your document class as
-% instructed by your conference organizers. Only withtimes
-% and notimes can be used in proceedings created by EasyChair
-%
-% The available formats are 'letterpaper' and 'a4paper' with
-% the former being the default if omitted as in the example
-% below.
-%
-\documentclass[procedia]{easychair}
-%\documentclass[debug]{easychair}
-%\documentclass[verbose]{easychair}
-%\documentclass[notimes]{easychair}
-%\documentclass[withtimes]{easychair}
-%\documentclass[a4paper]{easychair}
-%\documentclass[letterpaper]{easychair}
+% Template for Elsevier CRC journal article
+% version 1.2 dated 09 May 2011
 
-% This provides the \BibTeX macro
-\usepackage{doc}
-\usepackage{makeidx}
+% This file (c) 2009-2011 Elsevier Ltd.  Modifications may be freely made,
+% provided the edited file is saved under a different name
 
-\usepackage{eqnarray,amsmath}
+% This file contains modifications for Procedia Computer Science
+
+% Changes since version 1.1
+% - added "procedia" option compliant with ecrc.sty version 1.2a
+%   (makes the layout approximately the same as the Word CRC template)
+% - added example for generating copyright line in abstract
+
+%-----------------------------------------------------------------------------------
+
+%% This template uses the elsarticle.cls document class and the extension package ecrc.sty
+%% For full documentation on usage of elsarticle.cls, consult the documentation "elsdoc.pdf"
+%% Further resources available at http://www.elsevier.com/latex
+
+%-----------------------------------------------------------------------------------
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                          %%
+%% Important note on usage                                  %%
+%% -----------------------                                  %%
+%% This file should normally be compiled with PDFLaTeX      %%
+%% Using standard LaTeX should work but may produce clashes %%
+%%                                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% The '3p' and 'times' class options of elsarticle are used for Elsevier CRC
+%% The 'procedia' option causes ecrc to approximate to the Word template
+\documentclass[3p,times,procedia]{elsarticle}
+\flushbottom
+
+%% The `ecrc' package must be called to make the CRC functionality available
+\usepackage{ecrc}
+\usepackage[bookmarks=false]{hyperref}
+    \hypersetup{colorlinks,
+      linkcolor=blue,
+      citecolor=blue,
+      urlcolor=blue}
+%\usepackage{amsmath}
 \usepackage{subfig}
-\usepackage{graphicx}
-\usepackage{amsfonts}
+
+%% The ecrc package defines commands needed for running heads and logos.
+%% For running heads, you can set the journal name, the volume, the starting page and the authors
+
+%% set the volume if you know. Otherwise `00'
+\volume{00}
+
+%% set the starting page if not 1
+\firstpage{1}
+
+%% Give the name of the journal
+\journalname{Procedia Computer Science}
+
+%% Give the author list to appear in the running head
+%% Example \runauth{C.V. Radhakrishnan et al.}
+\runauth{V. Sovrasov}
+
+%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
+%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
+%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
+%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo
+
+%% Give the abbreviation of the Journal.
+\jid{procs}
+
+%% Give a short journal name for the dummy logo (if needed)
+%\jnltitlelogo{Computer Science}
+
+%% Hereafter the template follows `elsarticle'.
+%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.
+
+%% Elsevier CRC generally uses a numbered reference style
+%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
+%% If using BibTeX, use the style file elsarticle-num.bst
+
+%% End of ecrc-specific commands
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% The amssymb package provides various useful mathematical symbols
+
 \usepackage{amssymb}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
 
-\setlength{\textfloatsep}{4pt plus 1.0pt minus 1.5pt}
-\setlength{\floatsep}{4pt plus 1.0pt minus 1.5pt}
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers after \end{frontmatter}.
+%% \usepackage{lineno}
 
-% In order to save space or manage large tables or figures in a
-% landcape-like text, you can use the rotating and pdflscape
-% packages. Uncomment the desired from the below.
-%
-% \usepackage{rotating}
-% \usepackage{pdflscape}
+%% natbib.sty is loaded by default. However, natbib options can be
+%% provided with \biboptions{...} command. Following options are
+%% valid:
 
-% If you plan on including some algorithm specification, we recommend
-% the below package. Read more details on the custom options of the
-% package documentation.
-%
-% \usepackage{algorithm2e}
+%%   round  -  round parentheses are used (default)
+%%   square -  square brackets are used   [option]
+%%   curly  -  curly braces are used      {option}
+%%   angle  -  angle brackets are used    <option>
+%%   semicolon  -  multiple citations separated by semi-colon
+%%   colon  - same as semicolon, an earlier confusion
+%%   comma  -  separated by comma
+%%   numbers-  selects numerical citations
+%%   super  -  numerical citations as superscripts
+%%   sort   -  sorts multiple citations according to order in ref. list
+%%   sort&compress   -  like sort, but also compresses numerical citations
+%%   compress - compresses without sorting
+%%
+%% \biboptions{authoryear}
 
-\def\procediaConference{99th Conference on Topics of
-  Superb Significance (COOL 2014)}
+% \biboptions{}
 
-%\makeindex
+% if you have landscape tables
+\usepackage[figuresright]{rotating}
+%\usepackage{harvard}
+% put your own definitions here:x
+%   \newcommand{\cZ}{\cal{Z}}
+%   \newtheorem{def}{Definition}[section]
+%   ...
 
-%% Front Matter
+% add words to TeX's hyphenation exception list
+%\hyphenation{author another created financial paper re-commend-ed Post-Script}
+
+% declarations for front matter
+
+\begin{document}
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for the associated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for the associated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for the associated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
 %%
-% Regular title as in the article class.
-%
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
+\dochead{7th International Young Scientist Conference on Computational Science}%
+%% Use \dochead if there is an article header, e.g. \dochead{Short communication}
+%% \dochead can also be used to include a conference title, if directed by the editors
+%% e.g. \dochead{17th International Conference on Dynamical Processes in Excited States of Solids}
+
 \title{Comparison of dimensionality reduction schemes for derivative-free global optimization
@@
 
-% \titlerunning{} has to be set to either the main title or its shorter
-% version for the running heads. When processed by
-% EasyChair, this command is mandatory: a document without \titlerunning
-% will be rejected by EasyChair
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{<author name>}
+%% \address[label1]{<address>}
+%% \address[label2]{<address>}
 
-\titlerunning{Dimension reduction schemes in derivative-free GO}
+\author[a]{Vladislav Sovrasov\corref{cor1}}
 
-% Authors are joined by \and. Their affiliations are given by \inst, which indexes into the list
-% defined using \institute
-%
-\author {
-Vladislav Sovrasov
-}
-
-% Institutes for affiliations are also joined by \and,
-\institute  {
-  Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
-  \email{sovrasov.vlad@gmail.com}\\
-}
-%  \authorrunning{} has to be set for the shorter version of the authors' names;
-% otherwise a warning will be rendered in the running heads. When processed by
-% EasyChair, this command is mandatory: a document without \authorrunning
-% will be rejected by EasyChair
-
-\authorrunning{Sovrasov V.}
-
-\begin{document}
-
-\maketitle
-
-\keywords{Global optimization, Dimension reduction, Derivative-free algorithms, Global search algorithms}
+\address[a]{Lobachevsky State University of Nizhni Novgorod, 23 Prospekt Gagarina (Gagarin Avenue), 603950 Nizhnij Novgorod, Russia}
 
@@
 
-%\setcounter{tocdepth}{2}
-%{\small
-%\tableofcontents}
+\begin{keyword}
+Global optimization \sep Dimension reduction \sep Derivative-free algorithms \sep Global search algorithms
 
-%\section{To mention}
-%
-%Processing in EasyChair - number of pages.
-%
-%Examples of how EasyChair processes papers. Caveats (replacement of EC
-%class, errors).
+%% keywords here, in the form: keyword \sep keyword
+
+%% PACS codes here, in the form: \PACS code \sep code
+
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+
+\end{keyword}
+\cortext[cor1]{Corresponding author. E-mail address: sovrasov.vlad@gmail.com}
+\end{frontmatter}
+
+%\correspondingauthor[*]{Corresponding author. Tel.: +0-000-000-0000 ; fax: +0-000-000-0000.}
+\email{sovrasov.vlad@gmail.com}
+
+%%
+%% Start line numbering here if you want
+%%
+% \linenumbers
+
+%% main text
+
+%\enlargethispage{-7mm}
 
@@
 %------------------------------------------------------------------------------
-% Index
-%\printindex
-%\label{sect:bib}
-\bibliographystyle{unsrt}
+
+%% References
+%%
+%% Following citation commands can be used in the body text:
+%% Usage of \cite is as follows:
+%%   \cite{key}         ==>>  [#]
+%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
+%%
+
+%The citation must be used in following style: \cite{article-minimal} \cite{article-full} \cite{article-crossref} \cite{whole-journal}.
+%% References with BibTeX database:
+
 \bibliography{ysc_paper}
+\bibliographystyle{elsarticle-harv}
 
-%------------------------------------------------------------------------------
 \end{document}
 
-% EOF
+%%
+%% End of file `procs-template.tex'.

### 2018-07-10 2018/YSC/slides/slides.tex
YSC: the final version of slides

@@
 %set pages numeration
-\setbeamertemplate{footline}[frame number]
+%\setbeamertemplate{footline}[frame number]
+%\setlength\abovecaptionskip{-3pt}
+
+\addtobeamertemplate{navigation symbols}{}{%
+    \usebeamerfont{footline}%
+    \usebeamercolor[fg]{footline}%
+    \hspace{1em}%
+    \insertframenumber
+}
 \setbeamertemplate{headline}{}
-\setlength\abovecaptionskip{-1pt}
 
@@
   \end{displaymath}
-  \(\varphi(y(x))\) is non-smooth and has multiple local extremums even if \(\varphi(y)\) is unimodal.
+  \(\varphi(y(x))\) is non-smooth and has multiple local and \textbf{global} extremums even if \(\varphi(y)\) is unimodal.
   The latter problem is caused by loss of the information about \(N\)-d neighborhood after the transformation to the \(1\)-d space.
@@
 \begin{frame}
-  \frametitle{Smooth evolvent and non-univalent evolvent}
-      %\centerline{\includegraphics[width=0.9\textwidth]{smooth.pdf}}
+  \frametitle{Non-univalent evolvent}
+    One can try to recover all preimages of \(y\in\mathbb{R}^N\) and make optimization method aware of their existence\footnote{R.G.
+Strongin.
+Numerical Methods in Multiextremal Problems (in Russian), 1978}.
+This allows reducing the effect of growing amount of local minimas after dimension reduction.
+    According to the theory of Peano-type curves, each \(N\)-d point could have up to \(2^N\) preimages. For large \(N\) such preimages mining would be expensive.
       \begin{figure}[ht]
-        %\vspace*{-0.5cm}
-        \subfloat{\raisebox{.02\textheight}{\includegraphics[width=.35\textwidth]{smooth.pdf} }}
-        \subfloat{{\includegraphics[width=.25\textwidth]{noninjective.pdf} }}
+        \includegraphics[width=.25\textwidth]{noninjective.pdf}
       \end{figure}
@@
   \frametitle{Shifted and rotated evolvents}
+  To create a fixed amount of preimages one can use a pre-defined set of different evolvents.
+These evolvents could be shifted or rotated versions of the original one.
+Set of shifted evolvents\footnote{Strongin, R.G., Gergel, V.P., Barkalov, K.A.
+Parallel methods for global optimization problem
+solving (in Russian), 2009} is theoretically proven to generate at least one pair of close preimages if images are close and it perform better than the set of rotated curves.
   \begin{figure}[ht]
@@
 \begin{frame}
-  \frametitle{Rotated evolvent}
-  \begin{columns}
-    \begin{column}{0.5\textwidth}
+  \frametitle{Smooth evolvent}
+  Smooth functions are more predictable for optimizer, so smooth approximation of the Peano-like \(y(x)\) curve could impove convergence rate \footnote{Goryachih, A.
+A class of smooth modification of space-filling curves for global optimization problems, NET 2016}.
+  \begin{figure}[ht]
+    \includegraphics[width=0.45\textwidth]{smooth.pdf}
+  \end{figure}
 
-    \end{column}
-    \begin{column}{0.5\textwidth}
-      \centerline{\includegraphics[width=0.9\textwidth]{rotated.pdf}}
-    \end{column}
-  \end{columns}
 \end{frame}
@@
     \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
-    \item[Step 2.] Compute the evolvent \(y(x)\) and the function \(\varphi(y(x))\).
+    \item[Step 2.] Compute the evolvent \(y(x^k)\) and the function \(\varphi(y(x^k))\).
     \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
@@
 \begin{frame}
-  \frametitle{Conclusion and future work}
-    Already done:
+  \frametitle{Conclusions}
+    %Already done:
     \begin{itemize}
-      \item
+      \item the smooth evolvent and the non-univalent one demonstrate the best result in the problems of small dimensionality and can be applied successfully in solving the problems with the computational costly objective functions.
+      \item the shifted evolvents introduce large overhead costs on the execution of the method due to the requirement to adding an auxiliary constraint.
+About 95\% of iterations are overhead to fight the auxiliary constraint.
+      \item rotated evolvents perform almost the same as the shifted ones but without overhead.
     \end{itemize}
-    Future work:
-    \begin{itemize}
-      \item
-    \end{itemize}
+    %Future work:
+    %\begin{itemize}
+    %  \item
+    %\end{itemize}
 \end{frame}

### 2018-07-11 2018/YSC/ysc_paper.tex
YSC: fix typo
YSC: add information about grant

@@
 
-\address[a]{Lobachevsky State University of Nizhni Novgorod, 23 Prospekt Gagarina (Gagarin Avenue), 603950 Nizhnij Novgorod, Russia}
+\address[a]{Lobachevsky State University of Nizhni Novgorod, 23 Prospekt Gagarina (Gagarin Avenue), 603950 Nizhni Novgorod, Russia}
 
@@
 \end{keyword}
-\cortext[cor1]{Corresponding author. E-mail address: sovrasov.vlad@gmail.com}
+\cortext[cor1]{Corresponding author.}
 \end{frontmatter}
@@
 \subfloat[$r=5.0$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS2d_same_r_opt_pt_op.pdf}}\label{fig:gkls2d_acc}}
-    \subfloat[Minimal
-$r$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS2d_opt_pt_op.pdf}}\label{fig:gkls2d_opt}}
+    \subfloat[Minimal $r$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS2d_opt_pt_op.pdf}}\label{fig:gkls2d_opt}}
     \caption{Operating characteristics on GKLS 2d Simple class}
@@
 
+\section*{Acknowledgements}
+
+The study was supported by the Russian Science Foundation, project No 16-11-10150.
 %------------------------------------------------------------------------------

### 2018-07-16 2018/LeGo/LeGo.tex
final before review
Review before translation

@@
 
-%\pdfmapfile{+txfonts.map}
+\pdfmapfile{+txfonts.map}
 
-%%%%%ÄÎÁÀÂÈË ÄËß ÐÓÑÑÊÎÃÎ ÒÅÊÑÒÀ
-%\usepackage[utf8x]{inputenc}
-%\usepackage[english,russian]{babel}
-%\usepackage{cmap}
+%%%%%ƒŒ¡¿¬»À ƒÀﬂ –”—— Œ√Œ “≈ —“¿
+\usepackage[utf8x]{inputenc}
+\usepackage[english,russian]{babel}
+\usepackage{cmap}
 %%%%%
@@
 
-\affil{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia}
+\affil{Lobachevsky State University of Nizhni Novgorod, Gagarin avenue 23, 
+Nizhni Novgorod, Russia, 603950}
 
@@
 \begin{abstract}
-In the present work, further development of an approach to constructing the test global 
+In the present work, further development of an approach to constructing test global 
 optimization problems with nonlinear constraints is considered. In the generated problems, the 
@@
 problems with a conditional minimum located at the boundary of the feasible domain and of 
-specifying the number of constraints active in the minimizer have been added. The 
+specifying the number of constraints active at the minimizer have been added. The 
 demonstration of the developed approach is given on the example of a known index method for 
-solving the complex multiextremal optimization problems with non-convex constraints.
+solving complex multiextremal optimization problems with non-convex constraints.
 \end{abstract}
@@
 may be unknown, i.e. these ones may be defined by an algorithm for computing the function 
-values in the search domain (so called ``black-box''-functions). It is supposed that even a single 
-computing of a problem function value may be a time-consuming operation since it is related to 
-the necessity of numerical modeling in the applied problems (see, for example, 
-\cite{Menniti2008,Kvasov2015,Modorskii2017}).
+values in the search domain (so called ``black-box''-functions).
 
-The evaluation of efficiency of the developed methods is one of the key problems in the 
-optimization theory and applications. Unfortunately, it is difficult to obtain any theoretical 
-estimates in many cases. As a result, the comparison of the methods is performed by carrying 
-out the computational experiments on solving some test optimization problems in most cases. In 
-order to obtain a reliable evaluation of the efficiency of the methods, the sets of test problems 
-should be diverse and representative enough. The problem of choice of the test problems has 
-been considered in a lot of works (see, for example, 
-\cite{Floudas1999,Gaviano2003,Ali2005,Addis2007}). Unfortunately, in many cases, the 
-proposed sets contain a small number of test problems, and it is difficult to obtain the problems 
-with desired properties. The most important drawback consists of the fact that the constraints 
-are absent in the proposed test problems as a rule (or the constraints are relatively simple: linear, 
-convex, etc.).
+\Russian
+????????
+?????????
+????????
+?????
+??????????????????
+???????????
+?
+?????????????
+????????
+??????????, ?.?.
+???????
+????????
+???????????????
+?????
+?????
+????????
+???????????
+????????
+?????????
+??????
+???????????????
+??????????.
+?????????
+??????????
+?????????, ???
+???????, ??????
+???
+???????????.
+????????, ?????
+?????????
+?????????
+?????
+???
+???????????
+????
+??????????
+??????????
+(????????
+????????
+?
+\cite{Grishagin1997,Grishagin2015}).
+Another generator (\textit{GKLS generator}) for the functions of arbitrary dimensionality has been proposed in \cite{Gaviano2003}.
+GKLS-generator ???
+???????
+????????
+???
+????????????
+?????????
+???????
+??????????
+???????????
+\cite{Paulavicius2014,Sergeyev2015,Grishagin2018}.
+
+??
+??????
+??????????
+GKLS ???
+?????????
+??????
+?
+??????????
+????????
+?????
+?
+?????????????
+(Emmental-type GKLS-based test problems, \cite{Sergeyev2017}.
+?
+??????
+??????????
+???????
+????????
+????????
+????????????????
+GKLS-???????, ?????????????
+????????
+exterior of a ball ?
+?????????
+????????
+?
+?????????.
+?????????????
+????
+???????
+???????
+?/???
+???????????
+??????????, ???
+????????
+???????????
+???????
+??????????.
 
@@
 \begin{itemize}
-\item The possibility to generate the problems with the conditional minimum point located at the 
+\item The possibility to generate the problems with the constrained minimizer located at the 
 boundary of the feasible domain.
-\item The possibility to specify the number of constraints active in the optimum point.
+\item The possibility to control the number of constraints active in the optimum point.
 \end{itemize}
@@
 \section{TEST PROBLEM CLASSES}
-A well-known approach to investigating and comparing the multiextremal optimization 
-algorithms is based on testing these methods by solving a set of problems, chosen randomly 
-from some specially designed class.
 
-A generator for random samples of two-dimensional test functions has been 
-described, for example, in \cite{Grishagin1997,Grishagin2015}. %more refs
-Another generator (\textit{GKLS generator}) for the functions of arbitrary dimensionality with 
-known properties (the number of local minima, the size of their domains of attraction, the global 
-minimizer, etc.) has been proposed in \cite{Gaviano2003}.  Application of this generator for 
-studying some unconstrained optimization algorithms has been described in \cite{Paulavicius2014,Sergeyev2015}.
+\Russian
 
@@
 
-The location of the conditional minimum at the boundary of the feasible domain is an important 
+The location of the minimizer at the boundary of the feasible domain is an important 
 property featuring the applied constrained optimization problems. 
 When generating the problems according to the scheme described in \cite{Gergel2017}, 
-the conditional minimum, in general, will be located inside the feasible domain. Therefore, the 
-options of shifting the conditional minimum to the boundary of the feasible domain as well of 
-specifying the number of the active constraints in this one have been added in the new version 
+the constrained minimum point, in general, will be located inside the feasible domain. Therefore, the 
+options of shifting the minimizer to the boundary of the feasible domain as well of 
+controlling the number of the active constraints in this one have been added in the new version 
 of the problem generator.
 
-The shifting of the conditional minimum to the boundary of the feasible domain is performed in 
+The shifting of the minimizer to the boundary of the feasible domain is performed in 
 the following way.
 \begin{enumerate}
-\item A problem with an arbitrary location of the constrained minimum $\overline{y}$ is 
+\item A problem with an arbitrary location of the constrained minimum point $\overline{y}$ is 
 generated;
@@
 computation costs would increase linearly with increasing the dimensionality of the problem.
+\Russian 
+??? ???? ??????????? ?? ????? ??? $2N\left\|b-a\right\|/h$ ?????????? ???????? ??????? ??????? ???????.
 
-In the case, when it is necessary to generate a problem with a predefined number of active 
-constraints $S$ in the minimizer, the rules are modified as follows.
+In the case, when it is necessary to generate a problem with a number of active 
+constraints in the minimizer not less than $S$, the rules are modified as follows.
 \begin{enumerate}
-\item A problem with an arbitrary location of a conditional minimum $\overline{y}$ is 
+\item A problem with an arbitrary location of a constrained minimizer $\overline{y}$ is 
 generated;
@@
 \item The number $K$ of the constraints active at the point $y^*$ is determined.
-\item If $K>S$, $K-S$ constraints are selected from the set of active constraints, and these ones 
-are set to be inactive by subtracting a positive parameter $q$ from the right hand side.
 \item If $K<S$, $S-K$ constraints are selected from the set of the inactive ones, and and these 
@@
 performed. This way, the objective function reaches its minimum at the boundary of the feasible 
-domain with the predefined number of active constraints. 
+domain with the number of active constraints not less than $S$. 
 \end{enumerate}
@@
 As an illustration, the level lines of the objective functions and the zero-level lines of three 
-constraints for the problem constructed on the base of functions from \cite{Grishagin2015} with the volume 
-fraction of the feasible domains $\Delta = 0.5$ are shown in Fig.~\ref{fig:VAG}. The feasible 
-domains are highlighted by green.
-
-Fig.~\ref{fig:VAG} (a,b,c,d) correspond to different methods of generating the test problems.
+constraints for the problem constructed on the base of functions from \cite{Grishagin2015} 
+are shown in Fig.~\ref{fig:VAG}. The feasible domains are highlighted by green. 
+Figures~\ref{fig:VAG} (a,b,c,d) correspond to different methods of generating the test problems.
 
@@
 \end{enumerate}
+
+\Russian
+???????, ???
+volume fraction of the feasible domain ???
+??????
+????
+?????
+?????
+?????
+$\Delta = 0.5$, ?
+???
+?????????
+??????
+?????
+??
+?????
+???
+$\Delta = 0.5$ (?.?.
+???
+?????????
+??????
+???????
+????
+?????
+??????????
+???????
+?????
+??????????).
 
@@
 
-Fig.~\ref{fig:VAG} (a,b,c,d) also shows the points of 425, 381, 289 and 188 trials, 
+Figures~\ref{fig:VAG} (a,b,c,d) also shows the points of 425, 381, 289 and 188 trials, 
 correspondingly, performed by the \textit{index method} for solving constrained global 
@@
 
-The index method has been proposed and developed in 
+The index method has been developed in 
 \cite{Strongin2000,Sergeyev2001,Barkalov2002}. The approach is based on a separate 
@@
 of the feasible domain;
-	\item defining the number of constraints active at the optimum point.
+	\item controlling the number of constraints active at the optimum point.
 \end{itemize}

### 2018-07-19 2018/LeGo/LeGo.tex
After review

@@
 
-\pdfmapfile{+txfonts.map}
+%\pdfmapfile{+txfonts.map}
 
-%%%%%ƒŒ¡¿¬»À ƒÀﬂ –”—— Œ√Œ “≈ —“¿
-\usepackage[utf8x]{inputenc}
-\usepackage[english,russian]{babel}
-\usepackage{cmap}
+%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
+%\usepackage[utf8x]{inputenc}
+%\usepackage[english,russian]{babel}
+%\usepackage{cmap}
 %%%%%
@@
 
-\Russian
-????????
-?????????
-????????
-?????
-??????????????????
-???????????
-?
-?????????????
-????????
-??????????, ?.?.
-???????
-????????
-???????????????
-?????
-?????
-????????
-???????????
-????????
-?????????
-??????
-???????????????
-??????????.
-?????????
-??????????
-?????????, ???
-???????, ??????
-???
-???????????.
-????????, ?????
-?????????
-?????????
-?????
-???
-???????????
-????
-??????????
-??????????
-(????????
-????????
-?
-\cite{Grishagin1997,Grishagin2015}).
+%\Russian
+%Проблема построения генераторов многоэкстремальных тестовых задач с невыпуклыми ограничениями является актуальной, т.к.
+стандартный подход к исследованию и сравнению алгоритмов многоэкстремальной оптимизации основан на решении серии тестовых задач.
+%Известные генераторы порождают, как правило, задачи без ограничений.
+Например, способ построения двумерных тестовых функций был предложен Гришагиным (описание доступно в \cite{Grishagin1997,Grishagin2015}).
 Another generator (\textit{GKLS generator}) for the functions of arbitrary dimensionality has been proposed in \cite{Gaviano2003}.
-GKLS-generator ???
-???????
-????????
-???
-????????????
-?????????
-???????
-??????????
-???????????
-\cite{Paulavicius2014,Sergeyev2015,Grishagin2018}.
+GKLS-generator был успешно применен при исследовании некоторых методов глобальной оптимизации \cite{Paulavicius2014,Sergeyev2015,Grishagin2018}.
 
-??
-??????
-??????????
-GKLS ???
-?????????
-??????
-?
-??????????
-????????
-?????
-?
-?????????????
-(Emmental-type GKLS-based test problems, \cite{Sergeyev2017}.
-?
-??????
-??????????
-???????
-????????
-????????
-????????????????
-GKLS-???????, ?????????????
-????????
-exterior of a ball ?
-?????????
-????????
-?
-?????????.
-?????????????
-????
-???????
-???????
-?/???
-???????????
-??????????, ???
-????????
-???????????
-???????
-??????????.
+%На основе генератора GKLS был предложен подход к построению задач с ограничениями (Emmental-type GKLS-based test problems, \cite{Sergeyev2017}.
+В данном генераторе в качестве целевой функции используется модифицированная GKLS-функция, ограничениями являются exterior of a ball с заданными центрами и радиусами.
+Использование иных целевых функций и/или ограничений недоступно, что является недостатком данного генератора.
+
+The problem of developing the generators of the multiextremal test problems with non-convex 
+constraints is a relevant one since standard approach to investigating and comparing the 
+global optimization algorithms is based on solving a series of test problems.
+Well known generators generate the problems without constraints as a rule. For example, a 
+method for constructing two-dimensional test functions has been proposed by Grishagin (its 
+description can be found in \cite{Grishagin1997,Grishagin2015}). Another generator 
+(\textit{GKLS generator}) for the functions of arbitrary dimensionality has been proposed in 
+\cite{Gaviano2003}.  GKLS generator has been successfully applied in the investigation of some global 
+optimization methods \cite{Paulavicius2014,Sergeyev2015,Grishagin2018}.
+Based on the GKLS generator, an approach to constructing the constrained problems has been 
+proposed (Emmental-type GKLS-based test problems \cite{Sergeyev2017}). In this generator, 
+the modified GKLS-function was used as the objective function, the constraints were the 
+exterior of the balls with given centers and radii. The use of other objective functions and/or 
+constraints was not allowed that was a disadvantage of this generator.
 
@@
 \section{TEST PROBLEM CLASSES}
-
-\Russian
 
@@
 computation costs would increase linearly with increasing the dimensionality of the problem.
-\Russian 
-??? ???? ??????????? ?? ????? ??? $2N\left\|b-a\right\|/h$ ?????????? ???????? ??????? ??????? ???????.
+%\Russian 
+At that, no more than $2N\left\|b-a\right\|/h$ computations of the objective function and 
+constraints values will be required.
 
@@
 
-\Russian
-???????, ???
-volume fraction of the feasible domain ???
-??????
-????
-?????
-?????
-?????
-$\Delta = 0.5$, ?
-???
-?????????
-??????
-?????
-??
-?????
-???
-$\Delta = 0.5$ (?.?.
-???
-?????????
-??????
-???????
-????
-?????
-??????????
-???????
-?????
-??????????).
+%\Russian
+Note that the volume fraction of the feasible domain for the first three problems will be equal to 
+$\Delta = 0.5$, and will be no more than $\Delta = 0.5$ for the last problem (because the 
+volume of feasible domain can decrease when generating the problems with controlled number 
+of active constraints).
 
@@
   \begin{tabular}[b]{cc}
-    \includegraphics[width=.44\linewidth]{1.png} & \includegraphics[width=.44\linewidth]{2.png}\\
+    \includegraphics[width=.42\linewidth]{1.png} & \includegraphics[width=.42\linewidth]{2.png}\\
     \small (a) & \small (b) \\
-    \includegraphics[width=.44\linewidth]{3.png} & \includegraphics[width=.44\linewidth]{4.png}\\
+    \includegraphics[width=.42\linewidth]{3.png} & \includegraphics[width=.42\linewidth]{4.png}\\
     \small (c) & \small (d) 
@@
 as a blue point.
-
 The index method has been developed in 
@@
 accounting for each constraint of the problem and is not related to the use of the penalty 
-functions. According to the rules of index method, every iteration includes a sequential 
-checking of fulfillment of the problem constraints at this point. The first occurrence of violation 
-of any constraint terminates the trial and initiates the transition to the next iteration. This allows: 
+functions.
+
+%\Russian
+The index objective function utilized in the method essentially differs from the classical penalty one 
+ since the computing of the initial objective function value outside the feasible 
+domain is not required in the index method. Moreover, it is not required to compute the values 
+of all constraints outside the feasible domain as well but until the occurrence of the first 
+violated one only whereas in the classical penalty problems all constraints along with the 
+objective function are computed at every point.
+This allows the index method: 
 (i) accounting for the information on each constraint separately and (ii) solving the problems, in 

### 2018-07-30 2018/LeGo/LeGo.tex
final

@@
 problems with a conditional minimum located at the boundary of the feasible domain and of 
-specifying the number of constraints active at the minimizer have been added. The 
+controlling the number of constraints active at the minimizer have been added. The 
 demonstration of the developed approach is given on the example of a known index method for 
@@
 
-In the present paper, the methods for generating the global optimization test problems with non-convex constraints
+In the present paper, the methods for generating global optimization test problems with non-convex constraints
 \begin{eqnarray}
@@
 Lipschitz condition
-\[ \left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq i \leq m+1. \]
+\[ \left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq i \leq m+1, \]
 with the Lipschitz constants unknown a priori. The analytical formulae of the problem functions 
-may be unknown, i.e. these ones may be defined by an algorithm for computing the function 
+may be unknown, i.e. they may be defined by an algorithm for computing the function 
 values in the search domain (so called ``black-box''-functions).
@@
 The problem of developing the generators of the multiextremal test problems with non-convex 
-constraints is a relevant one since standard approach to investigating and comparing the 
-global optimization algorithms is based on solving a series of test problems.
-Well known generators generate the problems without constraints as a rule. For example, a 
+constraints is relevant since classes of test problems are used usually for investigating and comparing global optimization algorithms.
+Well known generators construct the problems without constraints as a rule. For example, a 
 method for constructing two-dimensional test functions has been proposed by Grishagin (its 
@@
 (\textit{GKLS generator}) for the functions of arbitrary dimensionality has been proposed in 
-\cite{Gaviano2003}.  GKLS generator has been successfully applied in the investigation of some global 
+\cite{Gaviano2003}.  GKLS generator has been successfully applied in the investigation of several global 
 optimization methods \cite{Paulavicius2014,Sergeyev2015,Grishagin2018}.
@@
 boundary of the feasible domain.
-\item The possibility to control the number of constraints active in the optimum point.
+\item The possibility to control the number of constraints active at the optimum point.
 \end{itemize}
@@
 generate the test global optimization problems with $m$ constraints has been proposed in \cite{Gergel2017}.
-In this paper, the rules, which allow formulating the constrained global optimization problems have been proposed so that:
+In the mentioned paper the rules which allow formulating the constrained global optimization problems have been proposed so that:
 \begin{itemize}
@@
 	\item the global minimizer of the objective function without accounting for the 
-constraints would be out of the feasible domain.
+constraints would be out of the feasible domain;
+	\item the value of the gradient at the global solution of the constrained problems is zero.
 \end{itemize}
@@
 options of shifting the minimizer to the boundary of the feasible domain as well of 
-controlling the number of the active constraints in this one have been added in the new version 
+controlling the number of the active constraints at the global solution have been added in the new version 
 of the problem generator.
@@
 In the case, when it is necessary to generate a problem with a number of active 
-constraints in the minimizer not less than $S$, the rules are modified as follows.
+constraints at the minimizer not less than $S$, the rules are modified as follows.
 \begin{enumerate}
@@
 \item From the point $\overline{y}$, a coordinate search of a feasible point $y^*$, for which at 
-least one constraint is active i. e. there is such index $j, 1 \leq j\leq m$ that $\left|g_j(y^*)\right| 
-\leq \delta$ is performed with a predefined step $h$.
+least one constraint is active (i.e. there is such index $j, 1 \leq j\leq m$, that $\left|g_j(y^*)\right| 
+\leq \delta$) is performed with a predefined step $h$.
 \item A rough estimate $y^*$ obtained is refined by a local method. 
 \item The number $K$ of the constraints active at the point $y^*$ is determined.
-\item If $K<S$, $S-K$ constraints are selected from the set of the inactive ones, and and these 
+\item If $K<S$, $S-K$ constraints are selected from the set of the inactive ones, and these 
 ones are set to be the active by adding a positive parameter $q$ to the right hand side.

### 2018-08-20 2018/ITNT/ITNT2018.tex

@@
 This assumption is natural for many applied problems since the relative variations of the functions generally cannot exceed a certain threshold determined by the bounded energy of changes occurring in the system under study.
-Also, a standard assumption is that the execution of a single trial (the computing of the values of the objective function and constraints at a point in the search domain) is a time-consuming operation since it
-utilized the results of numerical modeling.
+Also, a standard assumption is that the execution of a single trial (the computing of the values of the objective function and constraints at a point in the search domain) is a time-consuming operation since it utilized the results of numerical modeling.
 
@@
 
+\bibitem{Kovartsev2017}
+Kovartsev A N 2017 \textit{Computer Optics} \textbf{41(1)} 118--125 (In Russ.)
+
+\bibitem{Golovashkin2016}
+Malysheva S A and Golovashkin D L 2016 \textit{Computer Optics} \textbf{40(2)} 179--187 (In Russ.)
+
+\bibitem{Golovashkin2017}
+Vorotnikova D G and Golovashkin D L 2017 \textit{Computer Optics} \textbf{41(1)} 134--138 (In Russ.)
+
 \bibitem{Strongin2017}

### 2018-09-05 2018/RSD/MultiEvolvents/RSD.tex
RSD: fix style issues claimed by Springer's editors

@@
 \usepackage{amssymb}
+\usepackage[misc]{ifsym}
+\usepackage[labelsep=period]{caption}
 % Used for displaying a sample figure. If possible, figure files should
@@
 %
-\title{Comparison of dimensionality reduction schemes for parallel global optimization
-algorithms
-\thanks{The study was supported by the Russian Science Foundation, project No 16-11-
-10150}}
+\title{Comparison of Dimensionality Reduction Schemes for Parallel Global Optimization
+Algorithms}
 %
-\titlerunning{Dimension reduction schemes in parallel GO}
+\titlerunning{Dimension Reduction Schemes in Parallel GO}
 % If the paper title is too long for the running head, you can set
@@
 %
-\author{Konstantin Barkalov \and
-Vladislav Sovrasov \and
+\author{Konstantin Barkalov (\Letter),
+Vladislav Sovrasov and
 Ilya Lebedev}
@@
 \email{sovrasov.vlad@gmail.com}\\
-\email{ilya.lebedev@itmm.unn.ru}\\
-\url{http://hpc-education.unn.ru/}
+\email{ilya.lebedev@itmm.unn.ru}
 }
@@
 \section{Methods of Dimension Reduction}
-\subsection{Single evolvent}
+\subsection{Single Evolvent}
 
@@
 %------------------------------------------------------------------------------
-\subsection{Shifted evolvents}
+\subsection{Shifted Evolvents}
 \label{sec:shifted}
@@
 %------------------------------------------------------------------------------
-\subsection{Rotated evolvents}
+\subsection{Rotated Evolvents}
 The application of the scheme for building the multiple evolvents (hereinafter called the shifted
@@
 %------------------------------------------------------------------------------
-\subsection{Non-Univalent evolvent}
+\subsection{Non-Univalent Evolvent}
 %\begin{Russian}
@@
 %------------------------------------------------------------------------------
-\subsection{Smooth evolvent}
+\subsection{Smooth Evolvent}
 
@@
 %------------------------------------------------------------------------------
-\subsection{Parallel algorithm exploiting a set of evolvents}
+\subsection{Parallel Algorithm Exploiting a Set of Evolvents}
 \label{sec:parallel_evolvents}
@@
 %------------------------------------------------------------------------------
-\subsection{Comparison of the sequential evolvents}
+\subsection{Comparison of the Sequential Evolvents}
 \label{sec:seq_comp}
@@
 %------------------------------------------------------------------------------
-\subsection{Parallel rotated evolvents}
+\subsection{Parallel Rotated Evolvents}
 \label{sec:results_parallel}
@@
 
+\section*{Acknowledgements}
+The study was supported by the Russian Science Foundation, project No 16-11-
+10150
+
 % ---- Bibliography ----

### 2018-09-12 2018/RSD/MultiEvolvents/slides/slides.tex
RSD: update slides

--- 2018-07-10 2018/YSC/slides/slides.tex
@@
 
-\title{Comparison of dimensionality reduction schemes for derivative-free global optimization algorithms}
-\author{\textbf{Vladislav~Sovrasov}}
+\title{Comparison of Dimensionality Reduction Schemes for Parallel Global Optimization Algorithms}
+\author{\textbf{Konstantin~Barkalov}, \textbf{Vladislav~Sovrasov} and \textbf{Ilya~Lebedev}}
 \institute{Lobachevsky State University of Nizhni Novgorod}
@@
   \begin{figure}[ht]
-    %\vspace*{-0.5cm}
     \subfloat{\raisebox{.0\textheight}{\includegraphics[width=.35\textwidth]{shifted.pdf} }}
@@
   \end{figure}
-%  \begin{columns}
-%    \begin{column}{0.5\textwidth}
-%
-%    \end{column}
-%    \begin{column}{0.5\textwidth}
-%      \centerline{\includegraphics[width=0.9\textwidth]{shifted.pdf}}
-%    \end{column}
-%  \end{columns}
 \end{frame}
@@
   \frametitle{Smooth evolvent}
-  Smooth functions are more predictable for optimizer, so smooth approximation of the Peano-like \(y(x)\) curve could impove convergence rate \footnote{Goryachih, A.
+  Smooth functions are more predictable for optimizer, so smooth approximation of the Peano-like \(y(x)\) curve could improve convergence rate \footnote{Goryachih, A.
 A class of smooth modification of space-filling curves for global optimization problems, NET 2016}.
@@
 \begin{frame}
-  \frametitle{Optimization method}
+  \frametitle{Basic parallel optimization method}
   Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
@@
   \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Parallel optimization method with multiple evolvents}
+
 \end{frame}
@@
 \begin{frame}
+  \frametitle{Results of applying the parallel algorithm}
+
+  \begin{table}
+    \centering
+    \caption{Averaged numbers of iterations executed by the parallel algorithm for solving the test
+  optimization problems}
+    \label{tab:iterations}
+    \begin{tabular}{cccccccc}
+      \cline{3-8}\noalign{\smallskip}
+      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & &
+  \multicolumn{2}{c}{$N=5$}   \\
+      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
+      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
+  \textit{Hard}  \\
+      \noalign{\smallskip}\hline
+      I &
+      \parbox{0.25\textwidth}{
+      \begin{center}
+      \textbf{1 cluster node}
+      \end{center}		}
+        & \textit{1} & 12167 & 25635 & & 20979 & 187353  \\
+      &  & \textit{32} & 328 & 1268  & &   898 & 12208 \\
+      \hline \noalign{\smallskip}
+  II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
+    & \textit{1} & 25312 & 11103 & & 1472 & 17009 \\
+  &   & \textit{32} & 64 &   913 & & 47 & 345 \\
+      \noalign{\smallskip}\hline	\noalign{\smallskip}
+  III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
+    & \textit{1}  & 810 & 4351 & & 868 & 5697  \\
+  & & \textit{32} & 34  & 112  & & 35  & 868 \\
+      \noalign{\smallskip}\hline
+    \end{tabular}
+  \end{table}
+
+\end{frame}
+
+\begin{frame}
   \frametitle{Conclusions}

### 2018-09-16 2018/RSD/MultiEvolvents/slides/slides.tex
RSD: minir changes in slides
RSD: update slides

@@
 \setmathfont{Latin Modern Math}
+\setlength{\tabcolsep}{1pt}
 
@@
     \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
-    \item[Step 2.] Compute the evolvent \(y(x^k)\) and the function \(\varphi(y(x^k))\).
+    \item[Step 2.] Compute the evolvent \(y(x^{k+j})\) and the function \(\varphi(y(x^{k+j}))\), \(j=\overline{1,p}\).
     \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
-    \item[Step 4.] Choose an interval \((x_{t-1}, x_t)\) with the greatest characteristic and
-    compute objective \(f(y(x))\) in the point chosen using the decision rule \(d\):
+    \item[Step 4.] Choose \(p\) intervals \((x_{t_j-1}, x_{t_j})\) with the greatest characteristics and
+    compute objective \(f(y(x^{k+j}))\) in points chosen using the decision rule \(d\):
     \begin{displaymath}
-      x^{k+1}=d(t)\in (x_{t-1}, x_t)
+      x^{k+1+j}=d(t)\in (x_{t_j-1}, x_{t_j}),\:j=\overline{1,p}
     \end{displaymath}
-    \item[Step 5.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
+    \item[Step 5.] If \(x_{t_j}-x_{t_j-1}<\varepsilon\) for one of \(j=\overline{1,p}\), stop the method.
   \end{enumerate}
@@
   \frametitle{Parallel optimization method with multiple evolvents}
-
+  Using the multiple mapping allows solving initial problem by parallel solving the problems
+  \[
+  \min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
+  \]
+  on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a
+  separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
+  particular processor are interpreted as the results of the trials in the rest problems (in the
+  corresponding points \(x^{k_1},\dots,x^{k_S})\). In this approach, a trial at the point \(x^k \in
+  [0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
+  operations:
+  \begin{enumerate}
+    \setlength{\itemindent}{.1in}
+    \item[Step 1.] Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
+    \item[Step 2.] Inform the rest of processors about the start of the trial execution at the point \( y^k\) (the
+    blocking of the point \(y^k\) ).
+    \item[Step 3.] Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
+    trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
+    \(x{}^{k_1} ,\dots,x{}^{k_s} \)
+    \item[Step 4.] Inform the rest of processors about the trial results at the point \(y^k\).
+  \end{enumerate}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Results of comparison}
+  \frametitle{Evolvents comparison}
   \begin{figure}[ht]
@@
   \end{figure}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Choice of evolvent for the parallel algorithm}
+  \begin{table}
+  \begin{center}
+  \caption{Averaged number of computations of \(g_0\) and of \(\varphi\) when solving the
+  problems from GKLS 3d Simple class using the shifted evolvent}
+    \begin{tabular}{|l|{c}|{c}|{c}|}
+      \hline
+    $L$ & $calc(g_0)$ & $calc(\varphi)$ & $\frac{calc(g_0)}{calc(\varphi)}$ ratio \\
+    \hline
+    2 & 96247.9  & 6840.14 & 14.07\\
+    \hline
+    3 & 153131.0 & 7702.82 & 19.88\\
+    \hline
+    \end{tabular}
+    \label{tab:shifted_g0}
+  \end{center}
+  \end{table}
 \end{frame}
@@
 \begin{frame}
+  \frametitle{Results of applying the parallel algorithm}
+
+  \begin{table}
+    \centering
+    \caption{Speedup of parallel computations executed by the parallel algorithm}
+    \label{tab:speedup}
+    \begin{tabular}{cccccccc}
+      \cline{3-8}\noalign{\smallskip}
+      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & &
+  \multicolumn{2}{c}{$N=5$}   \\
+      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
+      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
+  \textit{Hard}  \\
+      \noalign{\smallskip}\hline
+      I &
+      \parbox{0.25\textwidth}{
+      \begin{center}
+      \textbf{1 cluster node}
+      \end{center}		}
+      & \textit{1}   & 12167(10.58s) & 25635(22.26s) & & 20979(22.78s) & 187353(205.83s)  \\
+    &  & \textit{32} & 37.1(18.03) & 20.2(8.55)  & &  23.3(8.77) & 15.4(9.68) \\
+    \hline \noalign{\smallskip}
+  II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
+  & \textit{1} &        0.5(0.33) & 2.3(0.86)  & & 14.3(6.61) & 11.0(6.06) \\
+  &   & \textit{32} & 190.1(9.59) & 28.1(1.08) & & 446.4(19.79) & 543.0(43.60) \\
+    \noalign{\smallskip}\hline	\noalign{\smallskip}
+  III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
+  & \textit{1}    & 15.0(6.05)  & 5.9(2.36)   & & 24.2(17.56)  & 32.9(24.87)  \\
+  & & \textit{32} & 357.9(2.36) & 228.9(2.64) & & 582.8(20.96) & 793.0(33.89) \\
+      \noalign{\smallskip}\hline
+    \end{tabular}
+  \end{table}
+
+\end{frame}
+
+\begin{frame}
   \frametitle{Conclusions}
@@
       \item rotated evolvents perform almost the same as the shifted ones but without overhead.
+      \item parallel optimization method shows up to 33x speedup on hard \(5d\) problems when using a set of rotated evolvents.
     \end{itemize}
@@
 \begin{frame}{{}}
-  \frametitle{ }
+  \frametitle{Q\&A}
   \begin{center}
-    \Large{Q\&A}
-
-\vspace{1cm}
+    \Large{Contacts:}
+\vspace{0.5cm}
 

### 2018-09-23 2018/RSD/MultiEvolvents/slides/slides.tex
RSD: update slides

@@
 \begin{frame}
+  \frametitle{Parallel optimization method with multiple evolvents}
+  \begin{figure}[ht]
+    \includegraphics[width=0.95\textwidth]{evolvents_parallel.pdf}
+  \end{figure}
+\end{frame}
+
+\begin{frame}
   \frametitle{Test problems}
@@
   \frametitle{Choice of evolvent for the parallel algorithm}
+  \begin{itemize}
+    \item Smooth evolvent is too computational heavy.
+    \item Non-univalent evolvent generates large and unpredictable amount of preimages.
+    \item Shifted evolvent generates huge amount of auxiliary points to handle additional constraint.
+  \end{itemize}
   \begin{table}
@@
     \begin{itemize}
-      \item the smooth evolvent and the non-univalent one demonstrate the best result in the problems of small dimensionality and can be applied successfully in solving the problems with the computational costly objective functions.
-      \item the shifted evolvents introduce large overhead costs on the execution of the method due to the requirement to adding an auxiliary constraint.
+      \item The smooth evolvent and the non-univalent one demonstrate the best result in the problems of small dimensionality and can be applied successfully in solving the problems with the computational costly objective functions.
+      \item The shifted evolvents introduce large overhead costs on the execution of the method due to the requirement to adding an auxiliary constraint.
 About 95\% of iterations are overhead to fight the auxiliary constraint.
-      \item rotated evolvents perform almost the same as the shifted ones but without overhead.
-      \item parallel optimization method shows up to 33x speedup on hard \(5d\) problems when using a set of rotated evolvents.
+      \item Rotated evolvents perform almost the same as the shifted ones but without overhead.
+      \item Parallel optimization method shows up to 43x speedup on hard \(5d\) problems when using a set of rotated evolvents.
     \end{itemize}

### 2018-09-23 2018/RSD/MultiEvolvents/slides/slides.tex
RSD: fix numeration of slides
RSD: fixes in slides

@@
 \graphicspath{ {../pictures/}{img/} }
+
 %set pages numeration
-%\setbeamertemplate{footline}[frame number]
-%\setlength\abovecaptionskip{-3pt}
-
-\addtobeamertemplate{navigation symbols}{}{%
-    \usebeamerfont{footline}%
-    \usebeamercolor[fg]{footline}%
-    \hspace{1em}%
-    \insertframenumber
-}
-\setbeamertemplate{headline}{}
+\newcommand\numbered{\setbeamertemplate{footline}{%
+  \vspace{-10em}
+   \raisebox{5pt}{\makebox[\paperwidth]{%
+     \hfill\makebox[10pt]{%
+       \usebeamerfont{footline}\usebeamercolor[fg]{footline}
+       \insertframenumber}}}}}
+\newcommand\unnumbered{\setbeamertemplate{footline}{}}
 
@@
 \begin{document}
+\numbered
+{
+\unnumbered
 \begin{frame}[noframenumbering,plain]
@@
 \end{frame}
+}
 
@@
     \lbrace y\in \mathbb{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\} \nonumber \\
-    \min\{f(y): y\in D\}=\min\{f(y(x)): x\in [0,1]\} \nonumber
+    \min\{\varphi(y): y\in D\}=\min\{\varphi(y(x)): x\in [0,1]\} \nonumber
   \end{gather}
@@
     \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
-    \item[Step 2.] Compute the evolvent \(y(x^{k+j})\) and the function \(\varphi(y(x^{k+j}))\), \(j=\overline{1,p}\).
-    \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
-    \item[Step 4.] Choose \(p\) intervals \((x_{t_j-1}, x_{t_j})\) with the greatest characteristics and
-    compute objective \(f(y(x^{k+j}))\) in points chosen using the decision rule \(d\):
+    \item[Step 2.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
+    \item[Step 3.] Choose \(p\) intervals \((x_{t_j-1}, x_{t_j})\) with the greatest characteristics and
+    compute objective \(\varphi(y(x^{k+j}))\) in points chosen using the decision rule \(d\):
     \begin{displaymath}
@@
     \end{displaymath}
-    \item[Step 5.] If \(x_{t_j}-x_{t_j-1}<\varepsilon\) for one of \(j=\overline{1,p}\), stop the method.
+    \item[Step 4.] If \(x_{t_j}-x_{t_j-1}<\varepsilon\) for one of \(j=\overline{1,p}\), stop the method.
   \end{enumerate}
@@
   \]
-  on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a
+  on a set of intervals $[0,1]$ by the basic method. Each one-dimensional problem is solved on a
   separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
@@
 \end{frame}
-
+{
+\unnumbered
 \begin{frame}{{}}
@@
 \end{frame}
-
+}
 \end{document}

### 2018-09-24 2018/RSD/MultiEvolvents/slides/slides.tex
RSD: update slides

@@
 \title{Comparison of Dimensionality Reduction Schemes for Parallel Global Optimization Algorithms}
-\author{\textbf{Konstantin~Barkalov}, \textbf{Vladislav~Sovrasov} and \textbf{Ilya~Lebedev}}
+\author{\textbf{Konstantin~Barkalov}, \underline{\textbf{Vladislav~Sovrasov}} and \textbf{Ilya~Lebedev}}
 \institute{Lobachevsky State University of Nizhni Novgorod}
@@
 These evolvents could be shifted or rotated versions of the original one.
-Set of shifted evolvents\footnote{Strongin, R.G., Gergel, V.P., Barkalov, K.A.
-Parallel methods for global optimization problem
-solving (in Russian), 2009} is theoretically proven to generate at least one pair of close preimages if images are close and it perform better than the set of rotated curves.
+Set of shifted evolvents\footnote{Strongin, R.G.
+Algorithms for multi-extremal mathematical programming problems employing the set of joint space-filling curves, Journal of Global Optimization 2(4), 357--378, 1992} is theoretically proven to generate at least one pair of close preimages if their images are close and it performs better than the a set of rotated curves in that sense.
   \begin{figure}[ht]
@@
   \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Cluster environment}
+  \begin{center}
+
+  The computational experiments have been carried out on the Lobachevsky supercomputer at  State University of Nizhni Novgorod.
+  One node includes 2 Intel Sandy Bridge E5-2660 2.2 GHz processors and 64 GB RAM.
+
+  Each node runs under CentOS 7 Linux with GCC 4.8 compiler and Intel MPI library.
+\end{center}
+
 \end{frame}

### 2018-11-02 2019/PAVT/pavt_paper.tex
Add pavt paper

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{svproc}
%
% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\usepackage{graphicx}
\usepackage{marvosym}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\usepackage{hyperref}
\def\UrlFont{\rmfamily}

\def\orcidID#1{\unskip$^{[#1]}$}
\def\letter{$^{\textrm{(\Letter)}}$}

\begin{document}
\mainmatter              % start of a contribution
%
\title{Comparison of several sequential and parallel derivative-free global optimization algorithms}
%
\titlerunning{Comparison of several optimization methods}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Vladislav Sovrasov\inst{1}\letter \and Semen Bevzuk\inst{1}}
%
\authorrunning{Vladislav Sovrasov et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Vladislav Sovrasov and Semen Bevzuk}
%
\institute{Lobachevsky State University of Nizhni Novgorod, Russia \\
  \email{sovrasov.vlad@gmail.com, semen.bevzuk@gmail.com}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
This work considers several stochastic and deterministic derivative-free global optimization algorithms. In the first part of the paper popular sequential open-source solvers are compared against the Globalizer solver, which is developed at the Lobachevsky State University. The Globalizer is designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows competitive performance with other similar solvers. The comparison is done on several sets of challenging multi-extremal benchmark functions. The second part of this work is devoted to comparison between the Globalizer and MIDACO solvers on systems with shared and distributed memory. MIDACO is a state-of-the-art global solver included to the TOMLAB optimization environment for MATLAB. Results of the benchmark show advantages of the Globalizer on small-dimensional, but sufficiently multi-extremal benchmark functions.
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{deterministic global optimization $\cdot$ stochastic global optimization
  $\cdot$ parallel numerical methods $\cdot$ derivative-free algorithms $\cdot$ black-box optimization}
\end{abstract}
%
\section{Introduction}
%
With this chapter, the preliminaries are over, and we begin the search
for periodic solutions to Hamiltonian systems. All this will be done in
the convex case; that is, we shall study the boundary-value problem
\begin{eqnarray*}
  \dot{x}&=&JH' (t,x)\\
  x(0) &=& x(T)
\end{eqnarray*}
with $H(t,\cdot)$ a convex function of $x$, going to $+\infty$ when
$\left\|x\right\| \to \infty$.

%
\subsection{Autonomous Systems}
%
In this section, we will consider the case when the Hamiltonian $H(x)$
is autonomous. For the sake of simplicity, we shall also assume that it
is $C^{1}$.

We shall first consider the question of nontriviality, within the
general framework of
$\left(A_{\infty},B_{\infty}\right)$-subquadratic Hamiltonians. In
the second subsection, we shall look into the special case when $H$ is
$\left(0,b_{\infty}\right)$-subquadratic,
and we shall try to derive additional information.
%
\subsubsection{The General Case: Nontriviality.}
%
We assume that $H$ is
$\left(A_{\infty},B_{\infty}\right)$-sub\-qua\-dra\-tic at infinity,
for some constant symmetric matrices $A_{\infty}$ and $B_{\infty}$,
with $B_{\infty}-A_{\infty}$ positive definite. Set:
\begin{eqnarray}
\gamma :&=&{\rm smallest\ eigenvalue\ of}\ \ B_{\infty} - A_{\infty} \\
  \lambda : &=& {\rm largest\ negative\ eigenvalue\ of}\ \
  J \frac{d}{dt} +A_{\infty}\ .
\end{eqnarray}

Theorem~\ref{ghou:pre} tells us that if $\lambda +\gamma < 0$, the
boundary-value problem:
\begin{equation}
\begin{array}{rcl}
  \dot{x}&=&JH' (x)\\
  x(0)&=&x (T)
\end{array}
\end{equation}
has at least one solution
$\overline{x}$, which is found by minimizing the dual
action functional:
\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt
\end{equation}
on the range of $\Lambda$, which is a subspace $R (\Lambda)_{L}^{2}$
with finite codimension. Here
\begin{equation}
  N(x) := H(x) - \frac{1}{2} \left(A_{\infty} x,x\right)
\end{equation}
is a convex function, and
\begin{equation}
  N(x) \le \frac{1}{2}
  \left(\left(B_{\infty} - A_{\infty}\right) x,x\right)
  + c\ \ \ \forall x\ .
\end{equation}

%
\begin{proposition}
Assume $H'(0)=0$ and $ H(0)=0$. Set:
\begin{equation}
  \delta := \liminf_{x\to 0} 2 N (x) \left\|x\right\|^{-2}\ .
  \label{eq:one}
\end{equation}

If $\gamma < - \lambda < \delta$,
the solution $\overline{u}$ is non-zero:
\begin{equation}
  \overline{x} (t) \ne 0\ \ \ \forall t\ .
\end{equation}
\end{proposition}
%
\begin{proof}
Condition (\ref{eq:one}) means that, for every
$\delta ' > \delta$, there is some $\varepsilon > 0$ such that
\begin{equation}
  \left\|x\right\| \le \varepsilon \Rightarrow N (x) \le
  \frac{\delta '}{2} \left\|x\right\|^{2}\ .
\end{equation}

It is an exercise in convex analysis, into which we shall not go, to
show that this implies that there is an $\eta > 0$ such that
\begin{equation}
  f\left\|x\right\| \le \eta
  \Rightarrow N^{\ast} (y) \le \frac{1}{2\delta '}
  \left\|y\right\|^{2}\ .
  \label{eq:two}
\end{equation}

\begin{figure}
\centering
\includegraphics[height=6.2cm]{figure}
\caption{This is the caption of the figure}
\label{fig:example}
\end{figure}

Since $u_{1}$ is a smooth function, we will have
$\left\|hu_{1}\right\|_\infty \le \eta$
for $h$ small enough, and inequality (\ref{eq:two}) will hold,
yielding thereby:
\begin{equation}
  \psi (hu_{1}) \le \frac{h^{2}}{2}
  \frac{1}{\lambda} \left\|u_{1} \right\|_{2}^{2} + \frac{h^{2}}{2}
  \frac{1}{\delta '} \left\|u_{1}\right\|^{2}\ .
\end{equation}

If we choose $\delta '$ close enough to $\delta$, the quantity
$\left(\frac{1}{\lambda} + \frac{1}{\delta '}\right)$
will be negative, and we end up with
\begin{equation}
  \psi (hu_{1}) < 0\ \ \ \ \ {\rm for}\ \ h\ne 0\ \ {\rm small}\ .
\end{equation}

On the other hand, we check directly that $\psi (0) = 0$. This shows
that 0 cannot be a minimizer of $\psi$, not even a local one.
So $\overline{u} \ne 0$ and
$\overline{u} \ne \Lambda_{o}^{-1} (0) = 0$. \qed
\end{proof}
%
\begin{corollary}
Assume $H$ is $C^{2}$ and
$\left(a_{\infty},b_{\infty}\right)$-subquadratic at infinity. Let
$\xi_{1},\allowbreak\dots,\allowbreak\xi_{N}$  be the
equilibria, that is, the solutions of $H' (\xi ) = 0$.
Denote by $\omega_{k}$
the smallest eigenvalue of $H'' \left(\xi_{k}\right)$, and set:
\begin{equation}
  \omega : = {\rm Min\,} \left\{\omega_{1},\dots,\omega_{k}\right\}\ .
\end{equation}
If:
\begin{equation}
  \frac{T}{2\pi} b_{\infty} <
  - E \left[- \frac{T}{2\pi}a_{\infty}\right] <
  \frac{T}{2\pi}\omega
  \label{eq:three}
\end{equation}
then minimization of $\psi$ yields a non-constant $T$-periodic solution
$\overline{x}$.
\end{corollary}
%

We recall once more that by the integer part $E [\alpha ]$ of
$\alpha \in \bbbr$, we mean the $a\in \bbbz$
such that $a< \alpha \le a+1$. For instance,
if we take $a_{\infty} = 0$, Corollary 2 tells
us that $\overline{x}$ exists and is
non-constant provided that:

\begin{equation}
  \frac{T}{2\pi} b_{\infty} < 1 < \frac{T}{2\pi}
\end{equation}
or
\begin{equation}
  T\in \left(\frac{2\pi}{\omega},\frac{2\pi}{b_{\infty}}\right)\ .
  \label{eq:four}
\end{equation}

%
\begin{proof}
The spectrum of $\Lambda$ is $\frac{2\pi}{T} \bbbz +a_{\infty}$. The
largest negative eigenvalue $\lambda$ is given by
$\frac{2\pi}{T}k_{o} +a_{\infty}$,
where
\begin{equation}
  \frac{2\pi}{T}k_{o} + a_{\infty} < 0
  \le \frac{2\pi}{T} (k_{o} +1) + a_{\infty}\ .
\end{equation}
Hence:
\begin{equation}
  k_{o} = E \left[- \frac{T}{2\pi} a_{\infty}\right] \ .
\end{equation}

The condition $\gamma < -\lambda < \delta$ now becomes:
\begin{equation}
  b_{\infty} - a_{\infty} <
  - \frac{2\pi}{T} k_{o} -a_{\infty} < \omega -a_{\infty}
\end{equation}
which is precisely condition (\ref{eq:three}).\qed
\end{proof}
%

\begin{lemma}
Assume that $H$ is $C^{2}$ on $\bbbr^{2n} \setminus \{ 0\}$ and
that $H'' (x)$ is non-de\-gen\-er\-ate for any $x\ne 0$. Then any local
minimizer $\widetilde{x}$ of $\psi$ has minimal period $T$.
\end{lemma}
%
\begin{proof}
We know that $\widetilde{x}$, or
$\widetilde{x} + \xi$ for some constant $\xi
\in \bbbr^{2n}$, is a $T$-periodic solution of the Hamiltonian system:
\begin{equation}
  \dot{x} = JH' (x)\ .
\end{equation}

There is no loss of generality in taking $\xi = 0$. So
$\psi (x) \ge \psi (\widetilde{x} )$
for all $\widetilde{x}$ in some neighbourhood of $x$ in
$W^{1,2} \left(\bbbr / T\bbbz ; \bbbr^{2n}\right)$.

But this index is precisely the index
$i_{T} (\widetilde{x} )$ of the $T$-periodic
solution $\widetilde{x}$ over the interval
$(0,T)$, as defined in Sect.~2.6. So
\begin{equation}
  i_{T} (\widetilde{x} ) = 0\ .
  \label{eq:five}
\end{equation}

Now if $\widetilde{x}$ has a lower period, $T/k$ say,
we would have, by Corollary 31:
\begin{equation}
  i_{T} (\widetilde{x} ) =
  i_{kT/k}(\widetilde{x} ) \ge
  ki_{T/k} (\widetilde{x} ) + k-1 \ge k-1 \ge 1\ .
\end{equation}

This would contradict (\ref{eq:five}), and thus cannot happen.\qed
\end{proof}
%
\paragraph{Notes and Comments.}
The results in this section are a
refined version of \cite{smit:wat};
the minimality result of Proposition
14 was the first of its kind.

To understand the nontriviality conditions, such as the one in formula
(\ref{eq:four}), one may think of a one-parameter family
$x_{T}$, $T\in \left(2\pi\omega^{-1}, 2\pi b_{\infty}^{-1}\right)$
of periodic solutions, $x_{T} (0) = x_{T} (T)$,
with $x_{T}$ going away to infinity when $T\to 2\pi \omega^{-1}$,
which is the period of the linearized system at 0.

\begin{table}
\caption{This is the example table taken out of {\it The
\TeX{}book,} p.\,246}
\begin{center}
\begin{tabular}{r@{\quad}rl}
\hline
\multicolumn{1}{l}{\rule{0pt}{12pt}
                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
\hline\rule{0pt}{12pt}
8000 B.C.  &     5,000,000& \\
  50 A.D.  &   200,000,000& \\
1650 A.D.  &   500,000,000& \\
1945 A.D.  & 2,300,000,000& \\
1980 A.D.  & 4,400,000,000& \\[2pt]
\hline
\end{tabular}
\end{center}
\end{table}
%
\begin{theorem} [Ghoussoub-Preiss]\label{ghou:pre}
Assume $H(t,x)$ is
$(0,\varepsilon )$-subquadratic at
infinity for all $\varepsilon > 0$, and $T$-periodic in $t$
\begin{equation}
  H (t,\cdot )\ \ \ \ \ {\rm is\ convex}\ \ \forall t
\end{equation}
\begin{equation}
  H (\cdot ,x)\ \ \ \ \ {\rm is}\ \ T{\rm -periodic}\ \ \forall x
\end{equation}
\begin{equation}
  H (t,x)\ge n\left(\left\|x\right\|\right)\ \ \ \ \
  {\rm with}\ \ n (s)s^{-1}\to \infty\ \ {\rm as}\ \ s\to \infty
\end{equation}
\begin{equation}
  \forall \varepsilon > 0\ ,\ \ \ \exists c\ :\
  H(t,x) \le \frac{\varepsilon}{2}\left\|x\right\|^{2} + c\ .
\end{equation}

Assume also that $H$ is $C^{2}$, and $H'' (t,x)$ is positive definite
everywhere. Then there is a sequence $x_{k}$, $k\in \bbbn$, of
$kT$-periodic solutions of the system
\begin{equation}
  \dot{x} = JH' (t,x)
\end{equation}
such that, for every $k\in \bbbn$, there is some $p_{o}\in\bbbn$ with:
\begin{equation}
  p\ge p_{o}\Rightarrow x_{pk} \ne x_{k}\ .
\end{equation}
\qed
\end{theorem}
%
\begin{example} [{{\rm External forcing}}]
Consider the system:
\begin{equation}
  \dot{x} = JH' (x) + f(t)
\end{equation}
where the Hamiltonian $H$ is
$\left(0,b_{\infty}\right)$-subquadratic, and the
forcing term is a distribution on the circle:
\begin{equation}
  f = \frac{d}{dt} F + f_{o}\ \ \ \ \
  {\rm with}\ \ F\in L^{2} \left(\bbbr / T\bbbz; \bbbr^{2n}\right)\ ,
\end{equation}
where $f_{o} : = T^{-1}\int_{o}^{T} f (t) dt$. For instance,
\begin{equation}
  f (t) = \sum_{k\in \bbbn} \delta_{k} \xi\ ,
\end{equation}
where $\delta_{k}$ is the Dirac mass at $t= k$ and
$\xi \in \bbbr^{2n}$ is a
constant, fits the prescription. This means that the system
$\dot{x} = JH' (x)$ is being excited by a
series of identical shocks at interval $T$.
\end{example}
%
\begin{definition}
Let $A_{\infty} (t)$ and $B_{\infty} (t)$ be symmetric
operators in $\bbbr^{2n}$, depending continuously on
$t\in [0,T]$, such that
$A_{\infty} (t) \le B_{\infty} (t)$ for all $t$.

A Borelian function
$H: [0,T]\times \bbbr^{2n} \to \bbbr$
is called
$\left(A_{\infty} ,B_{\infty}\right)$-{\it subquadratic at infinity}
if there exists a function $N(t,x)$ such that:
\begin{equation}
  H (t,x) = \frac{1}{2} \left(A_{\infty} (t) x,x\right) + N(t,x)
\end{equation}
\begin{equation}
  \forall t\ ,\ \ \ N(t,x)\ \ \ \ \
  {\rm is\ convex\ with\  respect\  to}\ \ x
\end{equation}
\begin{equation}
  N(t,x) \ge n\left(\left\|x\right\|\right)\ \ \ \ \
  {\rm with}\ \ n(s)s^{-1}\to +\infty\ \ {\rm as}\ \ s\to +\infty
\end{equation}
\begin{equation}
  \exists c\in \bbbr\ :\ \ \ H (t,x) \le
  \frac{1}{2} \left(B_{\infty} (t) x,x\right) + c\ \ \ \forall x\ .
\end{equation}

If $A_{\infty} (t) = a_{\infty} I$ and
$B_{\infty} (t) = b_{\infty} I$, with
$a_{\infty} \le b_{\infty} \in \bbbr$,
we shall say that $H$ is
$\left(a_{\infty},b_{\infty}\right)$-subquadratic
at infinity. As an example, the function
$\left\|x\right\|^{\alpha}$, with
$1\le \alpha < 2$, is $(0,\varepsilon )$-subquadratic at infinity
for every $\varepsilon > 0$. Similarly, the Hamiltonian
\begin{equation}
H (t,x) = \frac{1}{2} k \left\|k\right\|^{2} +\left\|x\right\|^{\alpha}
\end{equation}
is $(k,k+\varepsilon )$-subquadratic for every $\varepsilon > 0$.
Note that, if $k<0$, it is not convex.
\end{definition}
%

\paragraph{Notes and Comments.}
The first results on subharmonics were
obtained by Rabinowitz in \cite{fo:kes:nic:tue}, who showed the existence of
infinitely many subharmonics both in the subquadratic and superquadratic
case, with suitable growth conditions on $H'$. Again the duality
approach enabled Clarke and Ekeland in \cite{may:ehr:stein} to treat the
same problem in the convex-subquadratic case, with growth conditions on
$H$ only.

Recently, Michalek and Tarantello (see \cite{fost:kes} and \cite{czaj:fitz})
have obtained lower bound on the number of subharmonics of period $kT$,
based on symmetry considerations and on pinching estimates, as in
Sect.~5.2 of this article \cite{onlyurl}.

%
% ---- Bibliography ----
%
\bibliographystyle{spmpsci}
\bibliography{bibliography}{}

\end{document}

### 2018-11-22 2019/PAVT/pavt_paper.tex
PAVT: update paper

@@
 \usepackage{marvosym}
+\usepackage{amssymb}
 
@@
 \section{Introduction}
-%
-With this chapter, the preliminaries are over, and we begin the search
-for periodic solutions to Hamiltonian systems. All this will be done in
-the convex case; that is, we shall study the boundary-value problem
-\begin{eqnarray*}
-  \dot{x}&=&JH' (t,x)\\
-  x(0) &=& x(T)
-\end{eqnarray*}
-with $H(t,\cdot)$ a convex function of $x$, going to $+\infty$ when
-$\left\|x\right\| \to \infty$.
 
-%
-\subsection{Autonomous Systems}
-%
-In this section, we will consider the case when the Hamiltonian $H(x)$
-is autonomous. For the sake of simplicity, we shall also assume that it
-is $C^{1}$.
+\section{Related Work}
 
-We shall first consider the question of nontriviality, within the
-general framework of
-$\left(A_{\infty},B_{\infty}\right)$-subquadratic Hamiltonians. In
-the second subsection, we shall look into the special case when $H$ is
-$\left(0,b_{\infty}\right)$-subquadratic,
-and we shall try to derive additional information.
-%
-\subsubsection{The General Case: Nontriviality.}
-%
-We assume that $H$ is
-$\left(A_{\infty},B_{\infty}\right)$-sub\-qua\-dra\-tic at infinity,
-for some constant symmetric matrices $A_{\infty}$ and $B_{\infty}$,
-with $B_{\infty}-A_{\infty}$ positive definite. Set:
-\begin{eqnarray}
-\gamma :&=&{\rm smallest\ eigenvalue\ of}\ \ B_{\infty} - A_{\infty} \\
-  \lambda : &=& {\rm largest\ negative\ eigenvalue\ of}\ \
-  J \frac{d}{dt} +A_{\infty}\ .
-\end{eqnarray}
-
-Theorem~\ref{ghou:pre} tells us that if $\lambda +\gamma < 0$, the
-boundary-value problem:
+\section{Statement of Multidimensional Global Optimization Problem}
+In this paper, the core class of optimization problems, which can be solved using
+Globalizer, is formulated. This class involves the multidimensional global
+optimization problems without constraints, which can be defined in the following way:
 \begin{equation}
-\begin{array}{rcl}
-  \dot{x}&=&JH' (x)\\
-  x(0)&=&x (T)
+\label{eq:task}
+\begin{array}{cr}\\
+  \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
+  D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
 \end{array}
 \end{equation}
-has at least one solution
-$\overline{x}$, which is found by minimizing the dual
-action functional:
+with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function
+\(\varphi(y)\) satisfies the Lipschitz condition
 \begin{equation}
-  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
-  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt
+\label{eq:lip}
+|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
 \end{equation}
-on the range of $\Lambda$, which is a subspace $R (\Lambda)_{L}^{2}$
-with finite codimension. Here
+where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbb{R}^N\)
+space.
+\par
+Usually, the objective function \(\varphi(y)\) is defined as a computational procedure,
+according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
+(let us further call such a calculation \textit{a trial}). It is supposed that this procedure
+is time-consuming.
+\section{Dimension Reduction with Evolvents}
+Within the framework of the information-statistical global optimization theory,
+the Peano space-filling curves (or evolvents) \(y(x)\) mapping the interval \([0,1]\)
+onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
+reduction \cite{sergeyevStronginLera2013}, \cite{strongin1978},
+\cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
+\par
+As a result of the reduction, the initial multidimensional global optimization
+problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
 \begin{equation}
-  N(x) := H(x) - \frac{1}{2} \left(A_{\infty} x,x\right)
+\label{eq:oneDimTask}
+\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
 \end{equation}
-is a convex function, and
+\par
+It is important to note that this dimensionality reduction scheme transforms the % minimized
+Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
+function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.
 \begin{equation}
-  N(x) \le \frac{1}{2}
-  \left(\left(B_{\infty} - A_{\infty}\right) x,x\right)
-  + c\ \ \ \forall x\ .
+\label{eq:holder}
+|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
 \end{equation}
+where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
+constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem
+(\ref{eq:task}).
+\par
+The algorithms for the numerical construction of the Peano curve approximations are
+given in \cite{strSergGO}.
 
-%
-\begin{proposition}
-Assume $H'(0)=0$ and $ H(0)=0$. Set:
-\begin{equation}
-  \delta := \liminf_{x\to 0} 2 N (x) \left\|x\right\|^{-2}\ .
-  \label{eq:one}
-\end{equation}
+\par
+The computational scheme obtained as a result of the dimensionality reduction consists of the
+following:
+\begin{itemize}
+  \item The optimization algorithm performs the minimization of the reduced one-dimensional
+  function \(\varphi(y(x))\) from (\ref{eq:oneDimTask}),
+  \item After determining the next trial point \(x\), a multidimensional image \(y\) is calculated by
+using the mapping \(y(x)\),
+  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the point
+\(y\in D\),
+  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
+\end{itemize}
+\subsection{Rotated Evolvents}
+To overcome complexity of the $S$-evolvent and to preserve the information on the nearness of the points in
+the $N$-dimensional space, one more scheme of building of the multiple mappings was proposed.
+The building of a set of Peano curves not by the shift along the main diagonal of the hypercube
+but by rotation of the evolvents around the coordinate origin is a distinctive feature of the
+proposed scheme \cite{Gergel2009}.
+In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case
+$N=2$ are presented as an illustration.
+Taking into account the initial mapping, one can conclude that current implementation of the
+method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain
+onto the corresponding one-dimensional intervals. Moreover, the additional constraint  $g_0(y)
+\leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent. This
+method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to
+$2^N$) if necessary.
+\section{Sequential and Parallel Algorithm of Global Search}
+The optimization methods applied in Globalizer to solve the reduced problem
+(\ref{eq:oneDimTask}) are based on the MAGS method, which can be presented as follows ---
+see \cite{strongin1978}, \cite{strSergGO}.
+\par
+The initial iteration of the algorithm is performed at an arbitrary point \mbox{\(x^1\in(0,1)\)}.
+Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
+The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the
+following rules.
 
-If $\gamma < - \lambda < \delta$,
-the solution $\overline{u}$ is non-zero:
-\begin{equation}
-  \overline{x} (t) \ne 0\ \ \ \forall t\ .
-\end{equation}
-\end{proposition}
-%
-\begin{proof}
-Condition (\ref{eq:one}) means that, for every
-$\delta ' > \delta$, there is some $\varepsilon > 0$ such that
-\begin{equation}
-  \left\|x\right\| \le \varepsilon \Rightarrow N (x) \le
-  \frac{\delta '}{2} \left\|x\right\|^{2}\ .
-\end{equation}
+\textit{Rule 1}. Renumber the points of the preceding trials by the lower indices in order of
+increasing value of coordinates
+$0=x_0<x_1<...<x_{k+1}=1$.
 
-It is an exercise in convex analysis, into which we shall not go, to
-show that this implies that there is an $\eta > 0$ such that
-\begin{equation}
-  f\left\|x\right\| \le \eta
-  \Rightarrow N^{\ast} (y) \le \frac{1}{2\delta '}
-  \left\|y\right\|^{2}\ .
-  \label{eq:two}
-\end{equation}
+\textit{Rule 2}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq
+k+1\).
 
-\begin{figure}
-\centering
-\includegraphics[height=6.2cm]{figure}
-\caption{This is the caption of the figure}
-\label{fig:example}
-\end{figure}
+\textit{Rule 4}. Determine the interval with the maximum characteristic $R(t)=\max_{1\leq i
+\leq k+1}R(i)$.
 
-Since $u_{1}$ is a smooth function, we will have
-$\left\|hu_{1}\right\|_\infty \le \eta$
-for $h$ small enough, and inequality (\ref{eq:two}) will hold,
-yielding thereby:
-\begin{equation}
-  \psi (hu_{1}) \le \frac{h^{2}}{2}
-  \frac{1}{\lambda} \left\|u_{1} \right\|_{2}^{2} + \frac{h^{2}}{2}
-  \frac{1}{\delta '} \left\|u_{1}\right\|^{2}\ .
-\end{equation}
+\textit{Rule 5}. Execute a new trial at the point \(x^{k+1}\) located within the interval with the
+maximum characteristic from the previous step
+  $x^{k+1}=d(x_t)$.
 
-If we choose $\delta '$ close enough to $\delta$, the quantity
-$\left(\frac{1}{\lambda} + \frac{1}{\delta '}\right)$
-will be negative, and we end up with
-\begin{equation}
-  \psi (hu_{1}) < 0\ \ \ \ \ {\rm for}\ \ h\ne 0\ \ {\rm small}\ .
-\end{equation}
+The stopping condition, which terminated the trials, is defined by the inequality
+$\rho_t<\varepsilon$
+for the interval with the maximum characteristic from Step 4 and \(\varepsilon >0\) is the
+predefined
+accuracy of the optimization problem solution. If the stopping condition is not satisfied,
+the index \(k\) is incremented by 1, and the new global optimization iteration is executed.
 
-On the other hand, we check directly that $\psi (0) = 0$. This shows
-that 0 cannot be a minimizer of $\psi$, not even a local one.
-So $\overline{u} \ne 0$ and
-$\overline{u} \ne \Lambda_{o}^{-1} (0) = 0$. \qed
-\end{proof}
-%
-\begin{corollary}
-Assume $H$ is $C^{2}$ and
-$\left(a_{\infty},b_{\infty}\right)$-subquadratic at infinity. Let
-$\xi_{1},\allowbreak\dots,\allowbreak\xi_{N}$  be the
-equilibria, that is, the solutions of $H' (\xi ) = 0$.
-Denote by $\omega_{k}$
-the smallest eigenvalue of $H'' \left(\xi_{k}\right)$, and set:
-\begin{equation}
-  \omega : = {\rm Min\,} \left\{\omega_{1},\dots,\omega_{k}\right\}\ .
-\end{equation}
-If:
-\begin{equation}
-  \frac{T}{2\pi} b_{\infty} <
-  - E \left[- \frac{T}{2\pi}a_{\infty}\right] <
-  \frac{T}{2\pi}\omega
-  \label{eq:three}
-\end{equation}
-then minimization of $\psi$ yields a non-constant $T$-periodic solution
-$\overline{x}$.
-\end{corollary}
-%
+The convergence conditions and exact formulas for descision rules $R(i)$ and $d(x)$ of the
+described algorithm are given, for example, in \cite{strSergGO}.
 
-We recall once more that by the integer part $E [\alpha ]$ of
-$\alpha \in \bbbr$, we mean the $a\in \bbbz$
-such that $a< \alpha \le a+1$. For instance,
-if we take $a_{\infty} = 0$, Corollary 2 tells
-us that $\overline{x}$ exists and is
-non-constant provided that:
+\subsection{Parallel Algorithm Exploiting a Set of Evolvents}
+\label{sec:parallel_evolvents}
+Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the
+problems
+\[
+\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
+\]
+on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a
+separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
+particular processor are interpreted as the results of the trials in the rest problems (in the
+corresponding points \(x^{k_1},\dots,x^{k_S})\). In this approach, a trial at the point \(x^k \in
+[0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
+operations:
+\par
+1. Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
+\par
+2. Inform the rest of processors about the start of the trial execution at the point \( y^k\) (the
+blocking of the point \(y^k\) ).
+\par
+3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
+trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
+\(x{}^{k_1} ,\dots,x{}^{k_s} \)
+\par
+4. Inform the rest of processors about the trial results at the point \(y^k\).
+\par
+The decision rules for the mentioned parallel algorithm, in general, are the same as the rules of the
+sequential algorithm (except the method of the trial execution). Each processor has its own copy
+of the software realizing the computations of the problem functions and the decision rule of the
+index algorithm. For the organization of the interactions among the processors, the queues are
+created on each processor, where the processors store the information on the executed iterations
+in the form of the tuples: the processor number \(s\), the trial point \(x{}^{k_s}\).
+\par
+The mentioned parallelization scheme is implemented in the Globalizer system with the use of MPI technology. Main
+features of implementation consist in the following. A separate MPI-process is created for each
+of \(S\) one-dimensional problems being solved, usually, one process per one processor
+employed. Each process can use $p$ threads to parallel execute $p$ trials, usually one thread per an accessible core.
 
-\begin{equation}
-  \frac{T}{2\pi} b_{\infty} < 1 < \frac{T}{2\pi}
-\end{equation}
-or
-\begin{equation}
-  T\in \left(\frac{2\pi}{\omega},\frac{2\pi}{b_{\infty}}\right)\ .
-  \label{eq:four}
-\end{equation}
+\section{Review of Other Methods}
+\subsection{Sequential Methods}
+\begin{itemize}
+  \item Algorithm of global search (AGSl) (\url{https://github.com/sovrasov/ags_nlp_solver})
+  \item Multi Level Single Linkage (MLSL) (\url{https://nlopt.readthedocs.io/en/latest/NLopt_Algorithms/#mlsl-multi-level-single-linkage})
+  \item DIRECT (\url{https://nlopt.readthedocs.io/en/latest/NLopt_Algorithms/#direct-and-direct-l})
+  \item Locally-based DIRECT (DIRECT$l$) (\url{https://nlopt.readthedocs.io/en/latest/NLopt_Algorithms/#direct-and-direct-l})
+  \item Dual Simulated Annealing (\url{https://github.com/sgubianpm/sdaopt})
+  \item Differential Evolution (\url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.optimize.differential_evolution.html#scipy.optimize.differential_evolution})
+  \item Controlled Random Search (\url{https://nlopt.readthedocs.io/en/latest/NLopt_Algorithms/#controlled-random-search-crs-with-local-mutation})
+  \item Simple (\url{https://github.com/chrisstroemel/Simple})
+  \item StoGO (\url{https://nlopt.readthedocs.io/en/latest/NLopt_Algorithms/#stogo})
+\end{itemize}
+\subsection{MIDACO Parallel Solver}
+\section{Tools for Comparison of Global Optimization Algorithms}
+\section{Results of Comparison}
+\section{Conclusions}
+\section*{Acknowledgements}
+The study was supported by the Russian Science Foundation, project No 16-11-
+10150
 
-%
-\begin{proof}
-The spectrum of $\Lambda$ is $\frac{2\pi}{T} \bbbz +a_{\infty}$. The
-largest negative eigenvalue $\lambda$ is given by
-$\frac{2\pi}{T}k_{o} +a_{\infty}$,
-where
-\begin{equation}
-  \frac{2\pi}{T}k_{o} + a_{\infty} < 0
-  \le \frac{2\pi}{T} (k_{o} +1) + a_{\infty}\ .
-\end{equation}
-Hence:
-\begin{equation}
-  k_{o} = E \left[- \frac{T}{2\pi} a_{\infty}\right] \ .
-\end{equation}
-
-The condition $\gamma < -\lambda < \delta$ now becomes:
-\begin{equation}
-  b_{\infty} - a_{\infty} <
-  - \frac{2\pi}{T} k_{o} -a_{\infty} < \omega -a_{\infty}
-\end{equation}
-which is precisely condition (\ref{eq:three}).\qed
-\end{proof}
-%
-
-\begin{lemma}
-Assume that $H$ is $C^{2}$ on $\bbbr^{2n} \setminus \{ 0\}$ and
-that $H'' (x)$ is non-de\-gen\-er\-ate for any $x\ne 0$. Then any local
-minimizer $\widetilde{x}$ of $\psi$ has minimal period $T$.
-\end{lemma}
-%
-\begin{proof}
-We know that $\widetilde{x}$, or
-$\widetilde{x} + \xi$ for some constant $\xi
-\in \bbbr^{2n}$, is a $T$-periodic solution of the Hamiltonian system:
-\begin{equation}
-  \dot{x} = JH' (x)\ .
-\end{equation}
-
-There is no loss of generality in taking $\xi = 0$. So
-$\psi (x) \ge \psi (\widetilde{x} )$
-for all $\widetilde{x}$ in some neighbourhood of $x$ in
-$W^{1,2} \left(\bbbr / T\bbbz ; \bbbr^{2n}\right)$.
-
-But this index is precisely the index
-$i_{T} (\widetilde{x} )$ of the $T$-periodic
-solution $\widetilde{x}$ over the interval
-$(0,T)$, as defined in Sect.~2.6. So
-\begin{equation}
-  i_{T} (\widetilde{x} ) = 0\ .
-  \label{eq:five}
-\end{equation}
-
-Now if $\widetilde{x}$ has a lower period, $T/k$ say,
-we would have, by Corollary 31:
-\begin{equation}
-  i_{T} (\widetilde{x} ) =
-  i_{kT/k}(\widetilde{x} ) \ge
-  ki_{T/k} (\widetilde{x} ) + k-1 \ge k-1 \ge 1\ .
-\end{equation}
-
-This would contradict (\ref{eq:five}), and thus cannot happen.\qed
-\end{proof}
-%
-\paragraph{Notes and Comments.}
-The results in this section are a
-refined version of \cite{smit:wat};
-the minimality result of Proposition
-14 was the first of its kind.
-
-To understand the nontriviality conditions, such as the one in formula
-(\ref{eq:four}), one may think of a one-parameter family
-$x_{T}$, $T\in \left(2\pi\omega^{-1}, 2\pi b_{\infty}^{-1}\right)$
-of periodic solutions, $x_{T} (0) = x_{T} (T)$,
-with $x_{T}$ going away to infinity when $T\to 2\pi \omega^{-1}$,
-which is the period of the linearized system at 0.
-
-\begin{table}
-\caption{This is the example table taken out of {\it The
-\TeX{}book,} p.\,246}
-\begin{center}
-\begin{tabular}{r@{\quad}rl}
-\hline
-\multicolumn{1}{l}{\rule{0pt}{12pt}
-                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
-\hline\rule{0pt}{12pt}
-8000 B.C.  &     5,000,000& \\
-  50 A.D.  &   200,000,000& \\
-1650 A.D.  &   500,000,000& \\
-1945 A.D.  & 2,300,000,000& \\
-1980 A.D.  & 4,400,000,000& \\[2pt]
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-%
-\begin{theorem} [Ghoussoub-Preiss]\label{ghou:pre}
-Assume $H(t,x)$ is
-$(0,\varepsilon )$-subquadratic at
-infinity for all $\varepsilon > 0$, and $T$-periodic in $t$
-\begin{equation}
-  H (t,\cdot )\ \ \ \ \ {\rm is\ convex}\ \ \forall t
-\end{equation}
-\begin{equation}
-  H (\cdot ,x)\ \ \ \ \ {\rm is}\ \ T{\rm -periodic}\ \ \forall x
-\end{equation}
-\begin{equation}
-  H (t,x)\ge n\left(\left\|x\right\|\right)\ \ \ \ \
-  {\rm with}\ \ n (s)s^{-1}\to \infty\ \ {\rm as}\ \ s\to \infty
-\end{equation}
-\begin{equation}
-  \forall \varepsilon > 0\ ,\ \ \ \exists c\ :\
-  H(t,x) \le \frac{\varepsilon}{2}\left\|x\right\|^{2} + c\ .
-\end{equation}
-
-Assume also that $H$ is $C^{2}$, and $H'' (t,x)$ is positive definite
-everywhere. Then there is a sequence $x_{k}$, $k\in \bbbn$, of
-$kT$-periodic solutions of the system
-\begin{equation}
-  \dot{x} = JH' (t,x)
-\end{equation}
-such that, for every $k\in \bbbn$, there is some $p_{o}\in\bbbn$ with:
-\begin{equation}
-  p\ge p_{o}\Rightarrow x_{pk} \ne x_{k}\ .
-\end{equation}
-\qed
-\end{theorem}
-%
-\begin{example} [{{\rm External forcing}}]
-Consider the system:
-\begin{equation}
-  \dot{x} = JH' (x) + f(t)
-\end{equation}
-where the Hamiltonian $H$ is
-$\left(0,b_{\infty}\right)$-subquadratic, and the
-forcing term is a distribution on the circle:
-\begin{equation}
-  f = \frac{d}{dt} F + f_{o}\ \ \ \ \
-  {\rm with}\ \ F\in L^{2} \left(\bbbr / T\bbbz; \bbbr^{2n}\right)\ ,
-\end{equation}
-where $f_{o} : = T^{-1}\int_{o}^{T} f (t) dt$. For instance,
-\begin{equation}
-  f (t) = \sum_{k\in \bbbn} \delta_{k} \xi\ ,
-\end{equation}
-where $\delta_{k}$ is the Dirac mass at $t= k$ and
-$\xi \in \bbbr^{2n}$ is a
-constant, fits the prescription. This means that the system
-$\dot{x} = JH' (x)$ is being excited by a
-series of identical shocks at interval $T$.
-\end{example}
-%
-\begin{definition}
-Let $A_{\infty} (t)$ and $B_{\infty} (t)$ be symmetric
-operators in $\bbbr^{2n}$, depending continuously on
-$t\in [0,T]$, such that
-$A_{\infty} (t) \le B_{\infty} (t)$ for all $t$.
-
-A Borelian function
-$H: [0,T]\times \bbbr^{2n} \to \bbbr$
-is called
-$\left(A_{\infty} ,B_{\infty}\right)$-{\it subquadratic at infinity}
-if there exists a function $N(t,x)$ such that:
-\begin{equation}
-  H (t,x) = \frac{1}{2} \left(A_{\infty} (t) x,x\right) + N(t,x)
-\end{equation}
-\begin{equation}
-  \forall t\ ,\ \ \ N(t,x)\ \ \ \ \
-  {\rm is\ convex\ with\  respect\  to}\ \ x
-\end{equation}
-\begin{equation}
-  N(t,x) \ge n\left(\left\|x\right\|\right)\ \ \ \ \
-  {\rm with}\ \ n(s)s^{-1}\to +\infty\ \ {\rm as}\ \ s\to +\infty
-\end{equation}
-\begin{equation}
-  \exists c\in \bbbr\ :\ \ \ H (t,x) \le
-  \frac{1}{2} \left(B_{\infty} (t) x,x\right) + c\ \ \ \forall x\ .
-\end{equation}
-
-If $A_{\infty} (t) = a_{\infty} I$ and
-$B_{\infty} (t) = b_{\infty} I$, with
-$a_{\infty} \le b_{\infty} \in \bbbr$,
-we shall say that $H$ is
-$\left(a_{\infty},b_{\infty}\right)$-subquadratic
-at infinity. As an example, the function
-$\left\|x\right\|^{\alpha}$, with
-$1\le \alpha < 2$, is $(0,\varepsilon )$-subquadratic at infinity
-for every $\varepsilon > 0$. Similarly, the Hamiltonian
-\begin{equation}
-H (t,x) = \frac{1}{2} k \left\|k\right\|^{2} +\left\|x\right\|^{\alpha}
-\end{equation}
-is $(k,k+\varepsilon )$-subquadratic for every $\varepsilon > 0$.
-Note that, if $k<0$, it is not convex.
-\end{definition}
-%
-
-\paragraph{Notes and Comments.}
-The first results on subharmonics were
-obtained by Rabinowitz in \cite{fo:kes:nic:tue}, who showed the existence of
-infinitely many subharmonics both in the subquadratic and superquadratic
-case, with suitable growth conditions on $H'$. Again the duality
-approach enabled Clarke and Ekeland in \cite{may:ehr:stein} to treat the
-same problem in the convex-subquadratic case, with growth conditions on
-$H$ only.
-
-Recently, Michalek and Tarantello (see \cite{fost:kes} and \cite{czaj:fitz})
-have obtained lower bound on the number of subharmonics of period $kT$,
-based on symmetry considerations and on pinching estimates, as in
-Sect.~5.2 of this article \cite{onlyurl}.
-
-%
 % ---- Bibliography ----

### 2018-11-23 2019/PAVT/pavt_paper.tex
PAVT: finish intro

@@
 \usepackage{amssymb}
-
+\usepackage{cite}
+%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
+\usepackage[utf8x]{inputenc}
+\usepackage[english,russian]{babel}
+\usepackage{cmap}
+%%%%%
 % to typeset URLs, URIs, and DOIs
@@
 \section{Introduction}
+Нелинейная глобальная оптимизация невыпуклых функций традиционно считается одной из самых трудных
+задач математического программирования. Отыскание глобального минимума функции от нескольких переменных
+зачастую оказывается сложнее, чем локальная оптимизация в тысячемерном пространстве. Для последней может оказаться достаточно
+применения простейшего метода градиентного спуска, в то время как чтобы \textit{гаранитрованно} отыскать глобальный оптимум методам
+оптимизации приходится накапливать информацию о поведении целевой функции во всей области поиска \cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}. В последнее время стали популярны
+различные стохастические алгоритмы глобально оптимизации, прежде всего эволюционные \cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}.
+Они имеют довольно простую структуру, позволяют решать задачи большой размерности, но обеспечивают глобальную сходимость только в вероятностном смысле.
+
+В данной работе рассмотрены open-source реализации девяти различных методов глобальной оптимизации.
+Все алгоритмы были протестированы на наборе из 900 существенно многоэкстремальных функций, который был сгенерирован с
+помощью специализированных генераторов задач \cite{Gaviano2003, grishaginClass}.
+Помимо сравнения последовательных алгоритмов, на подмножестве из 200 тестовых задач произведено сравнение солвера MIDACO \cite{Schlueter2012} и программной системы Globalizer \cite{globalizerSystem} в условиях работы на суперкомпьютере Лобачевский.
 
@@
   \item Controlled Random Search (\url{https://nlopt.readthedocs.io/en/latest/NLopt_Algorithms/#controlled-random-search-crs-with-local-mutation})
-  \item Simple (\url{https://github.com/chrisstroemel/Simple})
   \item StoGO (\url{https://nlopt.readthedocs.io/en/latest/NLopt_Algorithms/#stogo})

### 2018-11-24 2019/PAVT/pavt_paper.tex
PAVT: update paper

@@
 \section{Introduction}
+
+\begin{Russian}
 Нелинейная глобальная оптимизация невыпуклых функций традиционно считается одной из самых трудных
@@
 Помимо сравнения последовательных алгоритмов, на подмножестве из 200 тестовых задач произведено сравнение солвера MIDACO \cite{Schlueter2012} и программной системы Globalizer \cite{globalizerSystem} в условиях работы на суперкомпьютере Лобачевский.
+\end{Russian}
 
@@
 
+\begin{Russian}
+Ранее в литературе рассмартивалось как сравнение между собой стохастических алгоритмов глобальной оптимизации \cite{Ali2005, JSSv060i06}, так и детерминированных \cite{posik2012, KVASOV2018245, Liberti2005}.
+В них большинство современных методов изучены довольно детально, но упор в основном делается на последовательные алгоритмы, а основными критериями эффективности метода оптимизации является его надёжность и скорость сходимости.
+В большинстве работ в качестве набора тестовых функций берётся набор известных тестовых задач (например, функция Растригина, Ackley function и др.).
+Размер такого набора обычно не превышает 100 различных функций, некоторые из которых могут быть одноэкстремальны (как функция Розенброка).
+
+В работе \cite{Beiranvand2017} софрмулированы некие общие принципы, которых, по мнению авторов, следует придерживаться при сравнении методов оптимизации.
+В частности, авторы говорят о преимуществах генераторов задач, позволяющих создать большие наборы задач, сводя случайные эфффекты при сравнении к минимуму.
+В то же время, использование одного генератора может оказаться недостаточно для исчерпывающего сравнения методов.
+Чтобы частично преодолеть эту проблему, авторы \cite{Beiranvand2017} советуют использовать несколько генераторов различной природы и создавать наборы задач различной сложности.
+
+Учитывая опыт предыдущих работ в области сравнения методов оптимизации, в данной работе будут использованы два генератора тестовых задач разной природы, с помощью которых сгенерировано 9 наборов по 100 задач разной сложности размерности от 2 до 5.
+Помимо сравнения последовательных методов, также в работе приводится сравнение эффективности двух параллельных алгоритмов.
+\end{Russian}
+
 \section{Statement of Multidimensional Global Optimization Problem}
 In this paper, the core class of optimization problems, which can be solved using
-Globalizer, is formulated. This class involves the multidimensional global
+global optimization methods, is formulated. This class involves the multidimensional global
 optimization problems without constraints, which can be defined in the following way:
@@
 is time-consuming.
-\section{Dimension Reduction with Evolvents}
+
+\section{Review of Considered Optimization Methods}
+
+\subsection{Parallel Algorithm of Global Search}
+\subsubsection{Dimension Reduction with Evolvents}
 Within the framework of the information-statistical global optimization theory,
@@
 onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
-reduction \cite{sergeyevStronginLera2013}, \cite{strongin1978},
-\cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
+reduction \cite{sergeyevStronginLera2013, strongin1978, strSergGO}.
 \par
@@
 \end{itemize}
-\subsection{Rotated Evolvents}
-To overcome complexity of the $S$-evolvent and to preserve the information on the nearness of the points in
-the $N$-dimensional space, one more scheme of building of the multiple mappings was proposed.
-The building of a set of Peano curves not by the shift along the main diagonal of the hypercube
-but by rotation of the evolvents around the coordinate origin is a distinctive feature of the
-proposed scheme \cite{Gergel2009}.
-In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case
-$N=2$ are presented as an illustration.
-Taking into account the initial mapping, one can conclude that current implementation of the
-method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain
-onto the corresponding one-dimensional intervals. Moreover, the additional constraint  $g_0(y)
-\leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent. This
-method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to
-$2^N$) if necessary.
-\section{Sequential and Parallel Algorithm of Global Search}
-The optimization methods applied in Globalizer to solve the reduced problem
+
+\subsubsection{Algorithm of Global Search on Shared Memory}
+\label{sub:ags}
+Parallel optimization methods applied in Globalizer to solve the reduced problem
 (\ref{eq:oneDimTask}) are based on the MAGS method, which can be presented as follows ---
@@
 
-\textit{Rule 4}. Determine the interval with the maximum characteristic $R(t)=\max_{1\leq i
-\leq k+1}R(i)$.
+\textit{Rule 3}. Determine the \(p\) intervals with the maximum characteristics $R(t_j)=\max_{1\leq i
+\leq k+1}R(i),\: j=\overline{1,p}$.
 
-\textit{Rule 5}. Execute a new trial at the point \(x^{k+1}\) located within the interval with the
-maximum characteristic from the previous step
-  $x^{k+1}=d(x_t)$.
+\textit{Rule 4}. Execute new trials at points \(x^{k+j},\: j=\overline{1,p}\) located within intervals with
+maximum characteristics from the previous step
+  $x^{k+j}=d(x_{t_j}),\: j=\overline{1,p}$.
 
 The stopping condition, which terminated the trials, is defined by the inequality
-$\rho_t<\varepsilon$
-for the interval with the maximum characteristic from Step 4 and \(\varepsilon >0\) is the
-predefined
-accuracy of the optimization problem solution. If the stopping condition is not satisfied,
-the index \(k\) is incremented by 1, and the new global optimization iteration is executed.
+$\rho_{t_j}<\varepsilon,\: j=\overline{1,p}$
+for the intervals with maximum characteristics from Step 3 and \(\varepsilon >0\) is the
+predefined accuracy of the optimization problem solution. If the stopping condition is not satisfied,
+the index \(k\) is incremented by \(p\), and the new global optimization iteration is executed.
+
+This method is employed in Globalizer to organize parallel computations on shared memory: each of \(p\)
+trials can be carried out on one of \(p\) local computation units.
 
@@
 
-\subsection{Parallel Algorithm Exploiting a Set of Evolvents}
-\label{sec:parallel_evolvents}
+\subsection{Parallel Algorithm on Distributed Memory Exploiting a Set of Evolvents}
+\subsubsection{Rotated Evolvents}
+One of the possible ways to overcome the negative effects of using a numerical
+approximation of evolvent (it destroys the information about the neighbor points in $\mathbb{R}^N$ space)
+consists in using the multiple mappings
+\begin{equation}
+  Y_L(x)=\left\{y^0(x),\ y^1(x),...,\ y^L(x)\right\}
+\end{equation}
+instead of single Peano curve $y(x)$ (see \cite{strSergGO}).
+The building of a set of Peano curves by rotation of the evolvents around the coordinate origin is a distinctive feature was proposed in \cite{Gergel2009}.
+Taking into account the initial mapping, one can conclude that current implementation of the
+method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain
+onto the corresponding one-dimensional intervals. This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to
+$2^N$) if necessary.
+
 Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the
@@
 \par
-The mentioned parallelization scheme is implemented in the Globalizer system with the use of MPI technology. Main
-features of implementation consist in the following. A separate MPI-process is created for each
+The mentioned parallelization scheme was proposed in \cite{Gergel2009} and implemented in the Globalizer system with the use of MPI technology. Main
+features of implementation consist in the following:
+\begin{itemize}
+  \item A separate MPI-process is created for each
  of \(S\) one-dimensional problems being solved, usually, one process per one processor
-employed. Each process can use $p$ threads to parallel execute $p$ trials, usually one thread per an accessible core.
+ employed.
+ \item Each process can use $p$ threads to parallel execute $p$ trials, usually one thread per an accessible core.
+\end{itemize}
 
-\section{Review of Other Methods}
+\subsection{MIDACO Parallel Solver}
+MIDACO (Mixed Integer Distributed Ant Colony Optimization) \cite{Schlueter2012} is a solver which implements the evolutionary algorithm
+the ant colony optimization metaheuristic for continuous search domains. ACO was originally
+proposed to solve the TSP, but later has been successfully adopted to solve mixed integer nonlinear programming problems \cite{SCHLUTER2009}.
+
+The MIDACO solver utilizes reverse communication architecture \cite{Schlueter2012}.
+Reverse communication means that the call of the objective function happens outside and independently of the MIDACO source code.
+Relying on this feature the following distributed parallelization scheme was implemented:
+\begin{enumerate}
+  \item MIDACO runs at the master node and generates \(n \times p\) new trial points at each iteration.
+  \item Each of \(n\) distributed computational nodes gets \(p\) points from the master and performs \(p\) parallel trials on local computing devices.
+  \item Each of \(n\) distributed computational nodes sends \(p\) values of the objective function to the master.
+\end{enumerate}
+
+In this scheme all distributed communications were implemented using the MPI library.
+Parallelism within a single node is powered by the OpenMP standard.
+
 \subsection{Sequential Methods}
 \begin{itemize}
-  \item Algorithm of global search (AGSl) (\url{https://github.com/sovrasov/ags_nlp_solver})
-  \item Multi Level Single Linkage (MLSL) (\url{https://nlopt.readthedocs.io/en/latest/NLopt_Algorithms/#mlsl-multi-level-single-linkage})
-  \item DIRECT (\url{https://nlopt.readthedocs.io/en/latest/NLopt_Algorithms/#direct-and-direct-l})
-  \item Locally-based DIRECT (DIRECT$l$) (\url{https://nlopt.readthedocs.io/en/latest/NLopt_Algorithms/#direct-and-direct-l})
-  \item Dual Simulated Annealing (\url{https://github.com/sgubianpm/sdaopt})
-  \item Differential Evolution (\url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.optimize.differential_evolution.html#scipy.optimize.differential_evolution})
-  \item Controlled Random Search (\url{https://nlopt.readthedocs.io/en/latest/NLopt_Algorithms/#controlled-random-search-crs-with-local-mutation})
-  \item StoGO (\url{https://nlopt.readthedocs.io/en/latest/NLopt_Algorithms/#stogo})
+  \item \textbf{Algorithm of Global Search}. Sequential version of the method described in Section \ref{sub:ags}.
+
+  \item \textbf{Locally-based Algorithm of Global Search (AGS\(l\))} \cite{indexMethod}. It's a modification of
+  the original AGS which make it more locally oriented by alternately using of two types of characteristics in the Rule 2 from the Section \ref{sub:ags}.
+
+  \item \textbf{Multi Level Single Linkage} \cite{Kan1987StochasticGO}. MLSL is an improved multistart algorithm.
+  It samples low-discrepancy starting points and does local optimizations from them. In contrast to the dummy multistart schemes
+  MLSL uses some clustering heuristics to avoid multiple local descents to already explored local minimas.
+
+  \item \textbf{DIRECT} \cite{Jones2009}.
+The algorithm is deterministic and recursively divides the search space and forms a tree of hyper-rectangles (boxes).
+DIRECT uses the objective function values and the Lipschitz condition (\ref{eq:lip}) to estimate promising boxes.
+
+  \item \textbf{Locally-based DIRECT (DIRECT$l$)} \cite{Gablonsky2001}. It's a variation of DIRECT which pays less attention to non-promising boxes and therefore
+  has less exploration power: it can converge faster on problems with few local minimas, but lost the global one in complicated cases.
+
+  \item \textbf{Dual Simulated Annealing} \cite{XIANG1997216}.
+This stochastic method is a combination of the Classical Simulated Annealing and the Fast Simulated Annealing coupled to a strategy for applying a local search on accepted locations.
+It converges much faster than both parent algorithms, CSA and FSA.
+
+  \item \textbf{Differential Evolution} \cite{Storn1997}. DE is an adaptation of the original genetic algorithm to
+  the continuous search domain.
+
+  \item \textbf{Controlled Random Search} \cite{Price1983}. The CRS starts with a set of random points and then defines
+  the next trial point in relation to a simplex chosen randomly from a stored configuration of points. CRS in not an
+  evolutional algorithm, although stores something like population and performs transformation resembling a mutation.
+
+  \item \textbf{StoGO} \cite{Madsen1998}. StoGO is dividing the search space into smaller hyper-rectangles via a branch-and-bound approach,
+  and searching them by a local-search algorithm, optionally including some randomness.
+
 \end{itemize}
-\subsection{MIDACO Parallel Solver}
+
+All the mentioned algorithms (except of AGS\(l\) are available in source codes as parts of wide-spread optimization pacages.
+AGS, DIRECT, DIRECT$l$, CRS, MLSL and StoGO are part of the NLOpt libray \cite{nlopt}. Differential Evolution and DSA can be found in
+the latest version of the SciPy \cite{scipy} package for Python.
+
 \section{Tools for Comparison of Global Optimization Algorithms}
+
 \section{Results of Comparison}
+
 \section{Conclusions}
+
 \section*{Acknowledgements}

### 2018-11-25 2019/PAVT/pavt_paper.tex
PAVT:  add description of test problems

@@
 %
-\author{Vladislav Sovrasov\inst{1}\letter \and Semen Bevzuk\inst{1}}
+\author{Vladislav Sovrasov\letter \and Semen Bevzuk}
 %
@@
 
-All the mentioned algorithms (except of AGS\(l\) are available in source codes as parts of wide-spread optimization pacages.
+All the mentioned algorithms (except of AGS\(l\) which implemented in the Globalizer system)
+are available in source codes as parts of wide-spread optimization packages.
 AGS, DIRECT, DIRECT$l$, CRS, MLSL and StoGO are part of the NLOpt libray \cite{nlopt}. Differential Evolution and DSA can be found in
@@
 \section{Tools for Comparison of Global Optimization Algorithms}
+\begin{Russian}
 
-\section{Results of Comparison}
+Использование сгенерированных некоторыми случайными механизмами
+наборов тестовых задач с известными решениями является одним из общепринятых подходов
+к сравнению алгоритмов оптимизации \cite{Beiranvand2017}. В данной работе
+будем использовать два генератора тестовых задач, порождающих задачи различной природы \cite{grishaginClass, Gaviano2003}.
+
+Обозначим набор задач, полученных с помощью первого генератора из \cite{grishaginClass} как \(F_{GR}\). Механизм построения задач \(F_{GR}\),
+не предусматривает контроль за сложностью задач и количеством локальных оптимумов, однако известно, что порождаемые функции
+являются существенно многоэкстремальными. Кроме того, задачи, порождаемые \(F_{GR}\) двухмерные. В данной работе будем использовать
+100 случайно сгенерированных функций из класса \(F_{GR}\).
+
+Генератор GKLS \cite{Gaviano2003} позволяет получать задачи заданной размерности и с заданным количеством экстремумов.
+Кроме того GKLS позволяет регулировать сложность задач, уменьшая или увеличивая размер области притяжения гломального минимума.
+В работе \cite{SergeyevKvasov2006} указаны параметры генератора, позволяющие породить наборы по 100 задач двух уровней сложности
+(Simple и Hard) размерностей 2, 3, 4 и 5. Следуя примеру авторов генератора GKLS, будем использовать предложенные ими параметры и,
+таким образом, добавим в тестовый ещё 800 задач различной размерности и сложности.
+
+Будем считать, что тестовая задача решена, если метод оптимизации провёл очередное испытание \(y^k\) в
+\(\delta\)-окрестности глобального минимума \(y^*\), т.е. $\left\|y^k-
+y^*\right\|\leqslant \delta = 0.01\left\|b-a\right\|$, где \(a\) и \(b\) --- левая и правая границы гиперкуба из (\ref{eq:task}).
+Если указанное соотношение не выполнено до истечения лимита на количество испытаний, то задача считается не решённой.
+Максимальный лимит на количество испыаний установлен для каждого класса задач, в соответствии с размерностью и сложностью (см. таблицу \ref{tab:limits}).
+
+\begin{table}
+\begin{center}
+\caption{Trials limit for each test problems class}
+  \begin{tabular}{|l|{c}|}
+    \hline
+  Class & Limit\\
+  \hline
+  \(F_{GR}\) & 5000 \\
+  \hline
+  GKLS 2d Simple & 8000 \\
+  \hline
+  GKLS 2d Hard & 9000 \\
+  \hline
+  GKLS 3d Simple & 15000 \\
+  \hline
+  GKLS 3d Hard & 25000 \\
+  \hline
+  GKLS 4d Simple & 150000 \\
+  \hline
+  GKLS 4d Hard & 250000 \\
+  \hline
+  GKLS 5d Simple & 350000 \\
+  \hline
+  GKLS 5d Hard & 600000 \\
+  \hline
+  \end{tabular}
+  \label{tab:limits}
+\end{center}
+\end{table}
+
+В качестве характеристик метода оптимизации на каждом из классов будем рассматривать среднее число
+испытаний, затраченное для решения одной задачи, и количество решённых задач. Чем меньше число испытаний, тем быстрее метод сходится
+к решению, а значит и меньше обращается к потенциально трудомкой процедуре вычислений целевой функции.
+Количество решённых задач гоорит о надёжности метода с заданными параметрами на решаемом классе тестовых задач.
+\end{Russian}
+
+\section{Results of Numerical Experiments}
+\subsection{Results of Sequential Algorithms}
+seq
+\begin{table}
+\begin{center}
+\caption{Trials limit for each test problems class}
+  \begin{tabular}{|l|{c}|{c}|{c}|}
+    \hline
+  Class & AGS & CRS & DE\\
+  \hline
+  \(F_{GR}\) & \(r=3\) & popsize=150 & mutation=(1.1,1.9), popsize=60 \\
+  \hline
+  GKLS 2d Simple & \(r=4.6\) & popsize=200 & mutation=(1.1,1.9), popsize=60 \\
+  \hline
+  GKLS 2d Hard & \(r=6.5\) & popsize=400 & mutation=(1.1,1.9), popsize=60 \\
+  \hline
+  GKLS 3d Simple & \(r=3.7\) & popsize=1000 & mutation=(1.1,1.9), popsize=70 \\
+  \hline
+  GKLS 3d Hard & \(r=4.4\) & popsize=2000 & mutation=(1.1,1.9), popsize=80 \\
+  \hline
+  GKLS 4d Simple & \(r=4.7\) & popsize=8000 & mutation=(1.1,1.9), popsize=90 \\
+  \hline
+  GKLS 4d Hard & \(r=4.9\) & popsize=16000 & mutation=(1.1,1.9), popsize=100 \\
+  \hline
+  GKLS 5d Simple & \(r=4\) & popsize=25000 & mutation=(1.1,1.9), popsize=120 \\
+  \hline
+  GKLS 5d Hard & \(r=4\) & popsize=30000 & mutation=(1.1,1.9), popsize=140 \\
+  \hline
+  \end{tabular}
+\end{center}
+\end{table}
+
+\subsection{Results of Parallel Algorithms}
+
+\begin{table}
+  \centering
+  \caption{Averaged numbers of iterations executed by the parallel algorithm for solving the test
+optimization problems}
+  \label{tab:iterations}
+  \begin{tabular}{cccccccc}
+    \cline{3-8}\noalign{\smallskip}
+    \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & &
+\multicolumn{2}{c}{$N=5$}   \\
+    \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
+    \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
+\textit{Hard}  \\
+    \noalign{\smallskip}\hline
+    I &
+    \parbox{0.25\textwidth}{
+    \begin{center}
+    \textbf{1 cluster node}
+    \end{center}		}
+      & \textit{1} & 12167 & 25635 & & 20979 & 187353  \\
+    &  & \textit{32} & 328 & 1268  & &   898 & 12208 \\
+    \hline \noalign{\smallskip}
+II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
+  & \textit{1} & 25312 & 11103 & & 1472 & 17009 \\
+&   & \textit{32} & 64 &   913 & & 47 & 345 \\
+    \noalign{\smallskip}\hline	\noalign{\smallskip}
+III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
+  & \textit{1}  & 810 & 4351 & & 868 & 5697  \\
+& & \textit{32} & 34  & 112  & & 35  & 868 \\
+    \noalign{\smallskip}\hline
+  \end{tabular}
+\end{table}
+
+\begin{table}
+  \centering
+  \caption{Speedup of parallel computations executed by the parallel algorithm}
+  \label{tab:speedup}
+  \begin{tabular}{cccccccc}
+    \cline{3-8}\noalign{\smallskip}
+    \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & &
+\multicolumn{2}{c}{$N=5$}   \\
+    \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
+    \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
+\textit{Hard}  \\
+    \noalign{\smallskip}\hline
+    I &
+    \parbox{0.25\textwidth}{
+    \begin{center}
+    \textbf{1 cluster node}
+    \end{center}		}
+    & \textit{1}   & 12167(10.58s) & 25635(22.26s) & & 20979(22.78s) & 187353(205.83s)  \\
+  &  & \textit{32} & 37.1(18.03) & 20.2(8.55)  & &  23.3(8.77) & 15.4(9.68) \\
+  \hline \noalign{\smallskip}
+II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
+& \textit{1} &        0.5(0.33) & 2.3(0.86)  & & 14.3(6.61) & 11.0(6.06) \\
+&   & \textit{32} & 190.1(9.59) & 28.1(1.08) & & 446.4(19.79) & 543.0(43.60) \\
+  \noalign{\smallskip}\hline	\noalign{\smallskip}
+III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
+& \textit{1}    & 15.0(6.05)  & 5.9(2.36)   & & 24.2(17.56)  & 32.9(24.87)  \\
+& & \textit{32} & 357.9(2.36) & 228.9(2.64) & & 582.8(20.96) & 793.0(33.89) \\
+    \noalign{\smallskip}\hline
+  \end{tabular}
+\end{table}
+
+\begin{Russian}
+
+\begin{table}
+\begin{center}
+\caption{Trials limit for each test problems class}
+  \begin{tabular}{|l|{c}|}
+    \hline
+  Class & Limit\\
+  \hline
+  \(F_{GR}\) & 5000 \\
+  \hline
+  GKLS 2d Simple & 8000 \\
+  \hline
+  GKLS 2d Hard & 9000 \\
+  \hline
+  GKLS 3d Simple & 15000 \\
+  \hline
+  GKLS 3d Hard & 25000 \\
+  \hline
+  GKLS 4d Simple & 150000 \\
+  \hline
+  GKLS 4d Hard & 250000 \\
+  \hline
+  GKLS 5d Simple & 350000 \\
+  \hline
+  GKLS 5d Hard & 600000 \\
+  \hline
+  \end{tabular}
+\end{center}
+\end{table}
+\end{Russian}
 

### 2018-11-25 2019/PAVT/pavt_paper.tex
PAVT:  finish section with experiments

@@
 \subsection{Parallel Algorithm on Distributed Memory Exploiting a Set of Evolvents}
+\label{sub:parallel_evolvents}
 \subsubsection{Rotated Evolvents}
@@
 \begin{center}
-\caption{Trials limit for each test problems class}
+\caption{Trials limit for the each of test problems classes}
   \begin{tabular}{|l|{c}|}
@@
 к решению, а значит и меньше обращается к потенциально трудомкой процедуре вычислений целевой функции.
-Количество решённых задач гоорит о надёжности метода с заданными параметрами на решаемом классе тестовых задач.
+Количество решённых задач говорит о надёжности метода с заданными параметрами на решаемом классе тестовых задач.
+Чтобы сделать величины, характеризующие надёжность и скорость сходимости, при подсчёте среднего количества испытаний учитывались тоолько решённые задачи.
 \end{Russian}
@@
 \subsection{Results of Sequential Algorithms}
-seq
+\begin{Russian}
+
+Результаты тех или иных алгоритмов на различных классах задач напрямую зависят от
+настроек алгоритмов. В большинстве случаев авторы программных реализаций ориентируются на
+задачи средней сложности и чтобы получить удовлетворительный резултьтат при решении
+существенно многоэкстремальных задач, требуется корректировка некоторых параметров.
+При проведении сравнения были заданы следующие параметры для методов:
+\begin{itemize}
+  \item в методе AGS\(l\) параметр чередования глобального и локального правил вычисления характеристик был задан равным 5:1;
+  \item в методах DIRECT и DIRECT\(l\) параметр \(\epsilon=10^{-4}\);
+  \item в методе SDA параметр \(visit=2.72\).
+\end{itemize}
+
+Остальные параметры варьировались в зависимости от класса задач (см. таблицу \ref{tab:params}).
+
 \begin{table}
 \begin{center}
-\caption{Trials limit for each test problems class}
+\caption{Class-specific parameters of optimization algorithms}
   \begin{tabular}{|l|{c}|{c}|{c}|}
     \hline
-  Class & AGS & CRS & DE\\
+    & AGS, AGS\(l\) & CRS & DE\\
   \hline
@@
   \end{tabular}
+  \label{tab:params}
+\end{center}
+\end{table}
+
+В таблицах \ref{tab:trials}, \ref{tab:solved} приведены результаты запуска методов оптимизации на рассматриваемых классах задач.
+Методы DIRECT и AGS\(l\) показали наилучшую скорость сходимости на всех классах, причём AGS\(l\) уступает
+DIRECT на задачах из классов Simple и обходит его на задачах классов Hard. Как видно из таблицы \ref{tab:solved},
+самыми надёжными методами являются детерминированные: AGS, AGS\(l\), DIRECT и DIRECT\(l\). Из стохастических методов
+самую высокую надёжность демонстрируют MLSL и SDA.
+
+\begin{table}
+\begin{center}
+\caption{Averaged number of trials executed by sequential methods for solving the test
+optimization problems}
+  \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
+    \hline
+    & AGS & AGS\(l\) & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
+  \hline
+  \(F_{GR}\)     & 193.1 &  \textbf{158.3} & 400.3 & 182.3 & 214.9 & 947.2 & 691.2 & 1257.3 & 1336.8 \\
+  \hline
+  GKLS 2d Simple &  254.9 & 217.6 & 510.6 & \textbf{189.0} & 255.2 & 556.8 & 356.3 & 952.2 & 1251.5 \\
+  \hline
+  GKLS 2d Hard   &  728.7 & \textbf{488.0} & 844.7 & 985.4 & 1126.7 & 1042.5 & 1637.9 & 1041.1 & 2532.2 \\
+  \hline
+  GKLS 3d Simple &  1372.1 & 1195.3 & 4145.8 & \textbf{973.6} & 1477.8 & 4609.2 & 2706.5 & 5956.94 & 3856.1 \\
+  \hline
+  GKLS 3d Hard   &  3636.1 & \textbf{1930.5} & 6787.0 & 2298.7 & 3553.3 & 5640.1 & 4708.4 & 6914.3 & 7843.2 \\
+  \hline
+  GKLS 4d Simple &  26654.1 & 11095.7 & 37436.8 & \textbf{7824.3} & 15994.1 & 41514.3 & 21417.9 & 19157.7 & 59895.4 \\
+  \hline
+  GKLS 4d Hard   &  54536.8 &  \textbf{23167.8} &  73779.3 &  23204.4 &  54489.9 &  80247.2 &  68815.5 &  27466.1 &  109328.1  \\
+  \hline
+  GKLS 5d Simple &  29810.0 & 11529.0 & 143575.0 & \textbf{7166.5} & 13970.5 & 52647.6 & 34255.3 & 73074.5 & 91580.4 \\
+  \hline
+  GKLS 5d Hard   &  113129.1 & 67652.7 & 165192.8 & \textbf{66327.4} & 164390.6 & 138766.2 & 116973.1 & 105496.9 & 155123.8 \\
+  \hline
+  \end{tabular}
+  \label{tab:trials}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Number of test optimization problems solved by sequential methods}
+  \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
+    \hline
+    & AGS & AGS\(l\) & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
+  \hline
+  \(F_{GR}\)     &  100 & 100 & 76  & 100 & 100 & 97  & 96  & 96  & 67\\
+  \hline
+  GKLS 2d Simple &  100 & 100 & 85  & 100 & 100 & 100 & 100 & 98  & 90\\
+  \hline
+  GKLS 2d Hard   &  100 & 100 & 74  & 100 & 100 & 100 & 93  & 85  & 77 \\
+  \hline
+  GKLS 3d Simple &  100 & 97  & 75  & 100 & 100 & 100 & 89  & 86  & 44 \\
+  \hline
+  GKLS 3d Hard   &  100  & 99   & 72   & 100  & 99   & 100  & 88   & 77   & 43 \\
+  \hline
+  GKLS 4d Simple &  100 & 100 & 46  & 100 & 100 & 94  & 78  & 59  & 16 \\
+  \hline
+  GKLS 4d Hard   &  100 & 100 & 47  & 99  & 97  & 94  & 72  & 32  & 10  \\
+  \hline
+  GKLS 5d Simple &  100 & 100 & 68  & 100 & 100 & 98  & 100 & 77  & 9  \\
+  \hline
+  GKLS 5d Hard   &  97  & 99  & 42  & 100 & 90  & 79  & 84  & 48  & 8 \\
+  \hline
+  \end{tabular}
+  \label{tab:solved}
 \end{center}
@@
 \subsection{Results of Parallel Algorithms}
+
+The computational experiments with parallel algorithms have been carried out on the Lobachevsky supercomputer at
+State University of Nizhni Novgorod. A computational node includes 2 Intel
+Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. Each CPU has 8 cores (i. e. total 16
+cores were available per a node). All computational nodes are working under CentOS Linux 7.2.
+All considered algorithms are implemented using C++ and compiled by GCC 4.8.5.
+
+Сравнение двух параллельных алгоритмов, реализованных в системах Globalizer и MIDACO
+было проведено на четырёхмерных задачах GKLS, т.к. задачи большей размерности не поддерживаются
+бесплатной версией MIDACO. Как и в случае с последовательными алгоритмами, для кадого класса задач
+были изменены параметры методов. Для MIDACO параметр \(focus = -1\). В случае с Globalizer параметр
+\(r\) был выбран следуя таблице \ref{tab:params}. С удвоением количества задействованных развёрток
+этот параметр уменьшался на 0.3, поскольку увеличение количества развёрток увеличивает и надёжность метода из секции \ref{sub:parallel_evolvents}.
+
+Если считать, что затраты на параллелизм пренебрежимо малы по сравнению с затратами на вычисление целевых функций в задачах оптимизации, то ускорение по времени от использования параллельного метода будет равно ускорению по итерациям.
+Однако, в действительности это предположение справедливо,
+тлолько если вычисление значения целевой функции достаточно трудоёмко. Во всех численных экспериментах время вычисления целевой функции занимает примерно $2\times 10^{-3}$с.
+
+In Table~\ref{tab:iterations}, an averaged number of iterations when solving 100 problems from
+each considered class is presented.
+The number of iterations is reduced considerably with increasing the number of nodes and the
+number of threads on each node. В скобках также указано количество решённых задач в каждом классе.
+Globalizer решает практически все задачи в обоих классах, в то время как MIDACO не решает многие задачи из класса GKLS 4d Hard.
+Стоит ещё отметить, что с ростом числа узлов надёжность MIDACO падает на обоих классах задач, в то время как надёжность Globalizer остаётся стабильной.
 
@@
     \cline{3-8}\noalign{\smallskip}
-    \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & &
-\multicolumn{2}{c}{$N=5$}   \\
+    \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{Globalizer} & &
+\multicolumn{2}{c}{MIDACO}   \\
     \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
@@
     \noalign{\smallskip}\hline
-    I &
-    \parbox{0.25\textwidth}{
-    \begin{center}
-    \textbf{1 cluster node}
-    \end{center}		}
-      & \textit{1} & 12167 & 25635 & & 20979 & 187353  \\
-    &  & \textit{32} & 328 & 1268  & &   898 & 12208 \\
+    I & \textbf{1 cluster node}
+      & \textit{1} &   25270 (100)  & 55180 (99) & & 27645 (98) & 72068 (71)  \\
+    &  & \textit{16} & 1765 (100)  & 3714 (100) & &  1640 (97) & 4304 (70) \\
     \hline \noalign{\smallskip}
-II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
-  & \textit{1} & 25312 & 11103 & & 1472 & 17009 \\
-&   & \textit{32} & 64 &   913 & & 47 & 345 \\
-    \noalign{\smallskip}\hline	\noalign{\smallskip}
-III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-  & \textit{1}  & 810 & 4351 & & 868 & 5697  \\
-& & \textit{32} & 34  & 112  & & 35  & 868 \\
+II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
+  & \textit{1} & 13056 (100) & 22938 (99) & & 10558 (89) & 27128 (73) \\
+&   & \textit{16} & 732 (100) & 1759 (100)  &  & 1130 (92) & 2254 (73) \\
+    \hline \noalign{\smallskip}
+III & \textbf{4 cluster nodes} %\multirow{3}{*}{}
+  & \textit{1}  & 5016 (100) & 12703 (100) & & 5777 (94) & 15980 (75) \\
+& & \textit{16} & 367 (100) & 776 (100) & & 529 (88) &  1264 (66) \\
+    \hline \noalign{\smallskip}
+VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
+  & \textit{1}  & 2103 (100) & 5063 (100) & & 2847 (97) & 9853 (83)\\
+& & \textit{16} & 145 (100)  & 310 (100)  & & 272 (83) & 774 (57)\\
+    \hline \noalign{\smallskip}
+V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
+  & \textit{1}  & 1155 (100) & 2399 (100) & & 2233 (98) & 6022 (86) \\
+& & \textit{16} & 76 (100)  & 159 (100)  & & 168 (87) & 393 (53)\\
     \noalign{\smallskip}\hline
@@
 \end{table}
+
+Среднее ускорение по итерациям и среднее ускорение по времени (в скобках) приведены в таблице \ref{tab:speedup}.
+В первой строке таблицы приведено среднее количество итераций и среднее время решения задачи в
+последовательном режиме (в скобках). MIDACO для при параллельных вычислениях использует
+меньшее количество коммуникаций, поэтому показатели ускорения монотонно возрастают с ростом количества
+вычислительных узлов. В Globalizer каждый узел должен переслать результаты проведения итерации
+всем другим узлам и с увеличением количества узлов и количества рабочих потоков на каждом узле
+эффективность распараллеливания падает. Это можно видеть в таблице \ref{tab:speedup}, когда при переходе от 4х
+узлов к 8 ускорение по времени при использовании 16 потоков упало. В случае одного потока на узел
+ускорение обоих методов оптимизации примерно одинаковое. Однако стоит заметить, что Globalizer выполняет
+меньше испытаний и при \(p=1\) тратит гораздо меньше времени на решение всех задач.
+С ростом количества узлов и потоков MIDACO догоняет Globalizer по среднему времени решения задачи,
+однако надёжность при этом падает.
+
+Аномальное сверхлинейное ускорение по времени
+MIDACO на классе GKLS 2d Hard при использовании 8 и 12 узлов можно объяснить тем, что, согласно таблице
+\ref{tab:iterations}, MIDACO решил больше задач в параллельном режиме, чем в последовательном, а значит, на
+меньшем количестве задач работал до исчерпания лимита испытаний.
 
@@
     \noalign{\smallskip}\hline
-    I &
-    \parbox{0.25\textwidth}{
-    \begin{center}
-    \textbf{1 cluster node}
-    \end{center}		}
-    & \textit{1}   & 12167(10.58s) & 25635(22.26s) & & 20979(22.78s) & 187353(205.83s)  \\
-  &  & \textit{32} & 37.1(18.03) & 20.2(8.55)  & &  23.3(8.77) & 15.4(9.68) \\
+I  & \textbf{1 cluster node}  %\multirow{3}{*}{}
+    & \textit{1}   & 25270 (27.3s) & 55180 (61.8s) & & 27645 (32.2s) & 72068 (132.5s)  \\
+  &  & \textit{16} & 14.3(11.7) & 14.9(12.6)  & &  16.9(14.4) & 16.7(14.4) \\
   \hline \noalign{\smallskip}
-II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
-& \textit{1} &        0.5(0.33) & 2.3(0.86)  & & 14.3(6.61) & 11.0(6.06) \\
-&   & \textit{32} & 190.1(9.59) & 28.1(1.08) & & 446.4(19.79) & 543.0(43.60) \\
+II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
+  & \textit{1}      &   1.9(1.9) & 2.4(2.2)  & & 2.6(1.7) & 2.7(2.3) \\
+  &   & \textit{16} & 34.5(20.4) & 31.4(18.8) & & 24.5(18.8) & 32.0(29.1) \\
     \noalign{\smallskip}\hline	\noalign{\smallskip}
-III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-& \textit{1}    & 15.0(6.05)  & 5.9(2.36)   & & 24.2(17.56)  & 32.9(24.87)  \\
-& & \textit{32} & 357.9(2.36) & 228.9(2.64) & & 582.8(20.96) & 793.0(33.89) \\
+III  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
+& \textit{1}      & 5.0(4.6) & 4.3(4.1) & & 4.8(3.9) & 4.5(4.4) \\
+&   & \textit{16} & 68.8(23.7) & 71.2(25.8) & & 52.3(32.9) & 57.0(50.2) \\
+  \noalign{\smallskip}\hline	\noalign{\smallskip}
+VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
+& \textit{1}    & 12.0(8.7) & 10.9(8.3) & & 9.7(8.2)    & 7.3(9.0)    \\
+& & \textit{16} & 174.1(4.7) & 177.8(5.3) & & 101.6(51.6) & 93.1(84.1) \\
+    \noalign{\smallskip}\hline
+    V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
+    & \textit{1}    & 21.9(11.4)  & 23.0(12.8)  & & 12.4(11.1)  & 12.0(14.8)  \\
+    & & \textit{16} & 333.5(2.7) & 347.9(3.0) & & 164.5(82.0) & 183.2(129.7) \\
         \noalign{\smallskip}\hline
@@
 
-\begin{Russian}
-
-\begin{table}
-\begin{center}
-\caption{Trials limit for each test problems class}
-  \begin{tabular}{|l|{c}|}
-    \hline
-  Class & Limit\\
-  \hline
-  \(F_{GR}\) & 5000 \\
-  \hline
-  GKLS 2d Simple & 8000 \\
-  \hline
-  GKLS 2d Hard & 9000 \\
-  \hline
-  GKLS 3d Simple & 15000 \\
-  \hline
-  GKLS 3d Hard & 25000 \\
-  \hline
-  GKLS 4d Simple & 150000 \\
-  \hline
-  GKLS 4d Hard & 250000 \\
-  \hline
-  GKLS 5d Simple & 350000 \\
-  \hline
-  GKLS 5d Hard & 600000 \\
-  \hline
-  \end{tabular}
-\end{center}
-\end{table}
-\end{Russian}
-
 \section{Conclusions}
 
+В работе были рассмотрены несколько последовательных и параллельных алгоритмов глобальной оптимизации.
+
+\end{Russian}
 \section*{Acknowledgements}

### 2018-11-25 2019/PAVT/pavt_paper.tex
PAVT:  finish section with conclusions

@@
 
-В работе были рассмотрены несколько последовательных и параллельных алгоритмов глобальной оптимизации.
+В статье были рассмотрены несколько последовательных и параллельных алгоритмов глобальной оптимизации.
+Было проведено сравнение эффективности их работы на множестве тестовых задач, по результатам которого можно сделать
+следующие выводы:
+\begin{itemize}
+  \item методы AGS\(l\), реализованный в системе Globalizer показал скорость сходимости и надёжность на уровне \(DIRECT\) и
+  превосходит многие другие агоритмы, реаизации которых есть в открытом доступе;
+  \item стохастические методы оптимизации проигрывают детерминированным по скорости сходимости и надёжности. Особенно
+  сильно это проявляется на более сложных многоэкстремальных задачах;
+  \item параллельная версия системы Globalizer демонстрирует хорошие показатели ускорения по времени при работе на нескольких узлах,
+  на каждом из которых вычисляется одно значение целевой за итерацию. При решении задач с быстро вычислимой челевой функцией и
+  использовании многопоточности на узлах показатели ускорения системы Globalizer деградируют с увеличением количества узлов;
+  \item система MIDACO больше всего подходит для несложных глобальных задач, целевые функции в которых могут быть быстро вычислимы.
+  В таком случае MIDACO достаточно надёжна и обеспечивает линейное ускорение с ростом количества узлов и параллельно выполняемых потоков на них.
+\end{itemize}
 
 \end{Russian}
+
 \section*{Acknowledgements}
 The study was supported by the Russian Science Foundation, project No 16-11-
-10150
+10150.
 

### 2018-11-26 2019/PAVT/pavt_paper.tex
PAVT: fix minor typos

@@
 различные стохастические алгоритмы глобально оптимизации, прежде всего эволюционные \cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}.
-Они имеют довольно простую структуру, позволяют решать задачи большой размерности, но обеспечивают глобальную сходимость только в вероятностном смысле.
+Они имеют довольно простую структуру, позволяют решать задачи большой размерности, однако обеспечивают глобальную сходимость только в вероятностном смысле.
 
@@
 
-В работе \cite{Beiranvand2017} софрмулированы некие общие принципы, которых, по мнению авторов, следует придерживаться при сравнении методов оптимизации.
-В частности, авторы говорят о преимуществах генераторов задач, позволяющих создать большие наборы задач, сводя случайные эфффекты при сравнении к минимуму.
+В работе \cite{Beiranvand2017} сформулированы некие общие принципы, которых, по мнению авторов, следует придерживаться при сравнении методов оптимизации.
+В частности, авторы говорят о преимуществах генераторов задач, позволяющих создать большие наборы задач, сводя случайные эффекты при сравнении к минимуму.
 В то же время, использование одного генератора может оказаться недостаточно для исчерпывающего сравнения методов.
@@
 Генератор GKLS \cite{Gaviano2003} позволяет получать задачи заданной размерности и с заданным количеством экстремумов.
-Кроме того GKLS позволяет регулировать сложность задач, уменьшая или увеличивая размер области притяжения гломального минимума.
+Кроме того GKLS позволяет регулировать сложность задач, уменьшая или увеличивая размер области притяжения глобального минимума.
 В работе \cite{SergeyevKvasov2006} указаны параметры генератора, позволяющие породить наборы по 100 задач двух уровней сложности
@@
 y^*\right\|\leqslant \delta = 0.01\left\|b-a\right\|$, где \(a\) и \(b\) --- левая и правая границы гиперкуба из (\ref{eq:task}).
-Если указанное соотношение не выполнено до истечения лимита на количество испытаний, то задача считается не решённой.
-Максимальный лимит на количество испыаний установлен для каждого класса задач, в соответствии с размерностью и сложностью (см. таблицу \ref{tab:limits}).
+Если указанное соотношение не выполнено до истечения лимита на количество испытаний, то задача считается нерешённой.
+Максимальный лимит на количество испытаний установлен для каждого класса задач, в соответствии с размерностью и сложностью (см. таблицу \ref{tab:limits}).
 
@@
 испытаний, затраченное для решения одной задачи, и количество решённых задач. Чем меньше число испытаний, тем быстрее метод сходится
-к решению, а значит и меньше обращается к потенциально трудомкой процедуре вычислений целевой функции.
+к решению, а значит и меньше обращается к потенциально трудоёмкой процедуре вычислений целевой функции.
 Количество решённых задач говорит о надёжности метода с заданными параметрами на решаемом классе тестовых задач.
-Чтобы сделать величины, характеризующие надёжность и скорость сходимости, при подсчёте среднего количества испытаний учитывались тоолько решённые задачи.
+Чтобы сделать величины, характеризующие надёжность и скорость сходимости, при подсчёте среднего количества испытаний учитывались только решённые задачи.
 \end{Russian}
@@
 настроек алгоритмов. В большинстве случаев авторы программных реализаций ориентируются на
-задачи средней сложности и чтобы получить удовлетворительный резултьтат при решении
+задачи средней сложности и чтобы получить удовлетворительный результат при решении
 существенно многоэкстремальных задач, требуется корректировка некоторых параметров.
@@
   \item методы AGS\(l\), реализованный в системе Globalizer показал скорость сходимости и надёжность на уровне \(DIRECT\) и
-  превосходит многие другие агоритмы, реаизации которых есть в открытом доступе;
+  превосходит многие другие алгоритмы, реализации которых есть в открытом доступе;
   \item стохастические методы оптимизации проигрывают детерминированным по скорости сходимости и надёжности. Особенно
@@
   \item параллельная версия системы Globalizer демонстрирует хорошие показатели ускорения по времени при работе на нескольких узлах,
-  на каждом из которых вычисляется одно значение целевой за итерацию. При решении задач с быстро вычислимой челевой функцией и
+  на каждом из которых вычисляется одно значение целевой за итерацию. При решении задач с быстро вычислимой целевой функцией и
   использовании многопоточности на узлах показатели ускорения системы Globalizer деградируют с увеличением количества узлов;

### 2018-11-26 2019/PAVT/pavt_paper.tex
PAVT:  add unrelated references, fix table size

@@
 помощью специализированных генераторов задач \cite{Gaviano2003, grishaginClass}.
-Помимо сравнения последовательных алгоритмов, на подмножестве из 200 тестовых задач произведено сравнение солвера MIDACO \cite{Schlueter2012} и программной системы Globalizer \cite{globalizerSystem} в условиях работы на суперкомпьютере Лобачевский.
+Помимо сравнения последовательных алгоритмов, на подмножестве из 200 тестовых задач произведено сравнение солвера MIDACO \cite{Schlueter2012} и программной системы Globalizer \cite{globalizerSystem,Strongin2018} в условиях работы на суперкомпьютере Лобачевский.
 \end{Russian}
@@
 in the form of the tuples: the processor number \(s\), the trial point \(x{}^{k_s}\).
+The modifications of the method taking into account multiply criteria are given in \cite{GERGEL2017,Gergel2018}.
 \par
@@
     \hline
-  Class & Limit\\
+  Problems class & Trials limit\\
   \hline
@@
 Количество решённых задач говорит о надёжности метода с заданными параметрами на решаемом классе тестовых задач.
-Чтобы сделать величины, характеризующие надёжность и скорость сходимости, при подсчёте среднего количества испытаний учитывались только решённые задачи.
+Чтобы сделать величины, характеризующие надёжность и скорость сходимости, независимыми при подсчёте среднего количества испытаний учитывались только решённые задачи.
 \end{Russian}
@@
 optimization problems}
+\resizebox{\textwidth}{!}{%
   \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
@@
   \hline
-  \end{tabular}
+\end{tabular}}
   \label{tab:trials}
@@
 \(r\) был выбран следуя таблице \ref{tab:params}. С удвоением количества задействованных развёрток
-этот параметр уменьшался на 0.3, поскольку увеличение количества развёрток увеличивает и надёжность метода из секции \ref{sub:parallel_evolvents}.
+этот параметр уменьшался на 0.3, поскольку увеличение количества развёрток увеличивает и надёжность метода из секции \ref{sub:parallel_evolvents}
+(подробнее см. \cite{SOVRASOV2018}).
 

### 2018-11-28 2019/PAVT/pavt_paper.tex
PAVT: initial english version

@@
-%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%% author.tex
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%% Springer
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %
-% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% RECOMMENDED
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
 %
@@
 \usepackage{cite}
-%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
-\usepackage[utf8x]{inputenc}
-\usepackage[english,russian]{babel}
-\usepackage{cmap}
-%%%%%
 % to typeset URLs, URIs, and DOIs
@@
 \begin{abstract}
-This work considers several stochastic and deterministic derivative-free global optimization algorithms.
-In the first part of the paper popular sequential open-source solvers are compared against the Globalizer solver, which is developed at the Lobachevsky State University.
-The Globalizer is designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows competitive performance with other similar solvers.
-The comparison is done on several sets of challenging multi-extremal benchmark functions.
-The second part of this work is devoted to comparison between the Globalizer and MIDACO solvers on systems with shared and distributed memory.
-MIDACO is a state-of-the-art global solver included to the TOMLAB optimization environment for MATLAB.
-Results of the benchmark show advantages of the Globalizer on small-dimensional, but sufficiently multi-extremal benchmark functions.
+This work considers several stochastic and deterministic derivative-free global optimization
+algorithms. In the first part of the paper popular sequential open-source solvers are compared against
+the Globalizer solver, which is developed at the Lobachevskii State University. The Globalizer is
+designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows
+competitive performance with other similar solvers. The comparison is done on several sets of
+challenging multi-extremal benchmark functions. The second part of this work is devoted to
+comparison between the Globalizer and MIDACO solvers on systems with shared and distributed
+memory. MIDACO is a state-of-the-art global solver included to the TOMLAB optimization
+environment for MATLAB. Results of the benchmark show advantages of the Globalizer on small-
+dimensional, but sufficiently multi-extremal benchmark functions.
 % We would like to encourage you to list your keywords within
@@
 \keywords{deterministic global optimization $\cdot$ stochastic global optimization
-  $\cdot$ parallel numerical methods $\cdot$ derivative-free algorithms $\cdot$ black-box optimization}
+  $\cdot$ parallel numerical methods $\cdot$ derivative-free algorithms $\cdot$ black-box
+optimization}
 \end{abstract}
@@
 
-\begin{Russian}
-Нелинейная глобальная оптимизация невыпуклых функций традиционно считается одной из самых трудных
-задач математического программирования. Отыскание глобального минимума функции от нескольких переменных
-зачастую оказывается сложнее, чем локальная оптимизация в тысячемерном пространстве. Для последней может оказаться достаточно
-применения простейшего метода градиентного спуска, в то время как чтобы \textit{гаранитрованно} отыскать глобальный оптимум методам
-оптимизации приходится накапливать информацию о поведении целевой функции во всей области поиска \cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}. В последнее время стали популярны
-различные стохастические алгоритмы глобально оптимизации, прежде всего эволюционные \cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}.
-Они имеют довольно простую структуру, позволяют решать задачи большой размерности, однако обеспечивают глобальную сходимость только в вероятностном смысле.
+The nonlinear global optimization of the non-convex functions is considered to be one of the most
+difficult problems of mathematical programming traditionally. The finding of the global minimum of
+a function of several variables appears to be more difficult than the local optimization in the
+thousand-dimensional space often. For the latter, the application of the simplest gradient descent
+method may appear to be enough whereas in order to \textit{guarantee} the finding of the global
+optimum by the optimization methods, one has to accumulate the information on the behavior of the
+objective function within the whole search domain
+\cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}. To date, various stochastic global
+optimization algorithms have attracted much attention, first of all, the evolution ones
+\cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}. These ones have rather simple structure
+and allow solving the problems of large dimensionality. However, the evolution algorithms ensure
+the global convergence in the probabilistic sense only.
 
-В данной работе рассмотрены open-source реализации девяти различных методов глобальной оптимизации.
-Все алгоритмы были протестированы на наборе из 900 существенно многоэкстремальных функций, который был сгенерирован с
-помощью специализированных генераторов задач \cite{Gaviano2003, grishaginClass}.
-Помимо сравнения последовательных алгоритмов, на подмножестве из 200 тестовых задач произведено сравнение солвера MIDACO \cite{Schlueter2012} и программной системы Globalizer \cite{globalizerSystem,Strongin2018} в условиях работы на суперкомпьютере Лобачевский.
-\end{Russian}
+In the present paper, the open source implementations of nine different global optimization methods
+were considered. All algorithms were tested using a set of 900 essentially multiextremal functions,
+which has been generated using special problem generators \cite{Gaviano2003, grishaginClass}.
+Besides the comparison of the sequential algorithms, a comparison of the MIDACO solver
+\cite{Schlueter2012} with the Globalizer software system \cite{globalizerSystem,Strongin2018} has been
+performed while running on Lobachevsky supercomputer using a subset of 200 test problems.
 
@@
 
-\begin{Russian}
-Ранее в литературе рассмартивалось как сравнение между собой стохастических алгоритмов глобальной оптимизации \cite{Ali2005, JSSv060i06}, так и детерминированных \cite{posik2012, KVASOV2018245, Liberti2005}.
-В них большинство современных методов изучены довольно детально, но упор в основном делается на последовательные алгоритмы, а основными критериями эффективности метода оптимизации является его надёжность и скорость сходимости.
-В большинстве работ в качестве набора тестовых функций берётся набор известных тестовых задач (например, функция Растригина, Ackley function и др.).
-Размер такого набора обычно не превышает 100 различных функций, некоторые из которых могут быть одноэкстремальны (как функция Розенброка).
+Earlier, the comparison of the stochastic global optimization algorithms \cite{Ali2005, JSSv060i06}
+as well as of the deterministic ones \cite{posik2012, KVASOV2018245, Liberti2005} between each
+other has been considered in the literature. In these works, the majority of modern methods have
+been studied in details, but the emphasis was made mainly on the sequential algorithms while the
+reliability and the convergence speed were the main criteria of efficiency of the optimization
+methods. In the majority of works, the sets of well-known test problems (for example, the Rastrigin
+functions, Ackley function, etc.) were taken as the sets of test functions. The sizes of such sets don't
+exceed 100 different functions usually, some of which can be the single-extremal ones (such as the
+Rosenbrock functions).
 
-В работе \cite{Beiranvand2017} сформулированы некие общие принципы, которых, по мнению авторов, следует придерживаться при сравнении методов оптимизации.
-В частности, авторы говорят о преимуществах генераторов задач, позволяющих создать большие наборы задач, сводя случайные эффекты при сравнении к минимуму.
-В то же время, использование одного генератора может оказаться недостаточно для исчерпывающего сравнения методов.
-Чтобы частично преодолеть эту проблему, авторы \cite{Beiranvand2017} советуют использовать несколько генераторов различной природы и создавать наборы задач различной сложности.
+In Ref. \cite{Beiranvand2017}, some general principles were formulated, which, in the author's
+opinion, should be obeyed when comparing the optimization methods. In particular, the authors say
+about the advantages of the problem generators allowing generating the large sets of problems thus
+minimizing the random effects when comparing the methods. At the same time, the use of a single
+generator can appear to be not enough for a comprehensive comparison of the methods. In order to
+overcome this problem in part, the authors of Ref. \cite{Beiranvand2017} advise to use several
+generators of various nature and to create the sets of problems of various complexity.
 
-Учитывая опыт предыдущих работ в области сравнения методов оптимизации, в данной работе будут использованы два генератора тестовых задач разной природы, с помощью которых сгенерировано 9 наборов по 100 задач разной сложности размерности от 2 до 5.
-Помимо сравнения последовательных методов, также в работе приводится сравнение эффективности двух параллельных алгоритмов.
-\end{Russian}
+Taking into account the experience of the preceding works in the field of comparison of the
+optimization methods, two generators of the test problems of different nature will be used in the
+present work. Using these ones, 9 sets of 100 problems of various complexity with the
+dimensionality varying from 2 to 5 were generated. Besides the comparison of the sequential
+methods, a comparison of the efficiencies of two parallel algorithms is presented in the work as well.
 
@@
 \(y\in D\),
-  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
+  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-
+dimensional function \(\varphi(y(x))\) at the point \(x\).
 \end{itemize}
@@
 
-\textit{Rule 4}. Execute new trials at points \(x^{k+j},\: j=\overline{1,p}\) located within intervals with
+\textit{Rule 4}. Execute new trials at points \(x^{k+j},\: j=\overline{1,p}\) located within intervals
+with
 maximum characteristics from the previous step
@@
 
-This method is employed in Globalizer to organize parallel computations on shared memory: each of \(p\)
+This method is employed in Globalizer to organize parallel computations on shared memory: each of
+\(p\)
 trials can be carried out on one of \(p\) local computation units.
 
-The convergence conditions and exact formulas for descision rules $R(i)$ and $d(x)$ of the
+The convergence conditions and exact formulas for decision rules $R(i)$ and $d(x)$ of the
 described algorithm are given, for example, in \cite{strSergGO}.
@@
 One of the possible ways to overcome the negative effects of using a numerical
-approximation of evolvent (it destroys the information about the neighbor points in $\mathbb{R}^N$ space)
+approximation of evolvent (it destroys the information about the neighbor points in
+$\mathbb{R}^N$ space)
 consists in using the multiple mappings
@@
 instead of single Peano curve $y(x)$ (see \cite{strSergGO}).
-The building of a set of Peano curves by rotation of the evolvents around the coordinate origin is a distinctive feature was proposed in \cite{Gergel2009}.
-Taking into account the initial mapping, one can conclude that current implementation of the
+The building of a set of Peano curves by rotation of the evolvents around the coordinate origin is a
+distinctive feature was proposed in \cite{Gergel2009}. Taking into account the initial mapping, one
+can conclude that current implementation of the
 method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain
-onto the corresponding one-dimensional intervals. This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to
+onto the corresponding one-dimensional intervals. This method for building a set of mappings can be
+''scaled'' easily to obtain more evolvents (up to
 $2^N$) if necessary.
@@
 corresponding points \(x^{k_1},\dots,x^{k_S})\). In this approach, a trial at the point \(x^k \in
-[0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
+[0,1]\) executed in the framework of the \(s^{th}\) problem, consists in the following sequence
+of
 operations:
@@
 \par
-3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
+3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and
+interpret the
 trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
@@
 \par
-The mentioned parallelization scheme was proposed in \cite{Gergel2009} and implemented in the Globalizer system with the use of MPI technology. Main
+The mentioned parallelization scheme was proposed in \cite{Gergel2009} and implemented in the
+Globalizer system with the use of MPI technology. Main
 features of implementation consist in the following:
@@
  employed.
- \item Each process can use $p$ threads to parallel execute $p$ trials, usually one thread per an accessible core.
+ \item Each process can use $p$ threads to parallel execute $p$ trials, usually one thread per an
+accessible core.
 \end{itemize}
@@
 \subsection{MIDACO Parallel Solver}
-MIDACO (Mixed Integer Distributed Ant Colony Optimization) \cite{Schlueter2012} is a solver which implements the evolutionary algorithm
+MIDACO (Mixed Integer Distributed Ant Colony Optimization) \cite{Schlueter2012} is a solver
+which implements the evolutionary algorithm
 the ant colony optimization metaheuristic for continuous search domains. ACO was originally
-proposed to solve the TSP, but later has been successfully adopted to solve mixed integer nonlinear programming problems \cite{SCHLUTER2009}.
+proposed to solve the TSP, but later has been successfully adopted to solve mixed integer nonlinear
+programming problems \cite{SCHLUTER2009}.
 
 The MIDACO solver utilizes reverse communication architecture \cite{Schlueter2012}.
-Reverse communication means that the call of the objective function happens outside and independently of the MIDACO source code.
+Reverse communication means that the call of the objective function happens outside and
+independently of the MIDACO source code.
 Relying on this feature the following distributed parallelization scheme was implemented:
 \begin{enumerate}
-  \item MIDACO runs at the master node and generates \(n \times p\) new trial points at each iteration.
-  \item Each of \(n\) distributed computational nodes gets \(p\) points from the master and performs \(p\) parallel trials on local computing devices.
-  \item Each of \(n\) distributed computational nodes sends \(p\) values of the objective function to the master.
+  \item MIDACO runs at the master node and generates \(n \times p\) new trial points at each
+iteration.
+  \item Each of \(n\) distributed computational nodes gets \(p\) points from the master and performs
+\(p\) parallel trials on local computing devices.
+  \item Each of \(n\) distributed computational nodes sends \(p\) values of the objective function to
+the master.
 \end{enumerate}
@@
 \begin{itemize}
-  \item \textbf{Algorithm of Global Search}. Sequential version of the method described in Section \ref{sub:ags}.
+  \item \textbf{Algorithm of Global Search}. Sequential version of the method described in Section
+\ref{sub:ags}.
 
-  \item \textbf{Locally-based Algorithm of Global Search (AGS\(l\))} \cite{indexMethod}. It's a modification of
-  the original AGS which make it more locally oriented by alternately using of two types of characteristics in the Rule 2 from the Section \ref{sub:ags}.
+  \item \textbf{Locally-based Algorithm of Global Search (AGS\(l\))} \cite{indexMethod}. It's a
+modification of
+  the original AGS which make it more locally oriented by alternately using of two types of
+characteristics in the Rule 2 from the Section \ref{sub:ags}.
 
-  \item \textbf{Multi Level Single Linkage} \cite{Kan1987StochasticGO}. MLSL is an improved multistart algorithm.
-  It samples low-discrepancy starting points and does local optimizations from them. In contrast to the dummy multistart schemes
-  MLSL uses some clustering heuristics to avoid multiple local descents to already explored local minimas.
+  \item \textbf{Multi Level Single Linkage} \cite{Kan1987StochasticGO}. MLSL is an improved
+multistart algorithm.
+  It samples low-discrepancy starting points and does local optimizations from them. In contrast to
+the dummy multistart schemes
+  MLSL uses some clustering heuristics to avoid multiple local descents to already explored local
+minima.
 
-  \item \textbf{DIRECT} \cite{Jones2009}.
-The algorithm is deterministic and recursively divides the search space and forms a tree of hyper-rectangles (boxes).
-DIRECT uses the objective function values and the Lipschitz condition (\ref{eq:lip}) to estimate promising boxes.
+  \item \textbf{DIRECT} \cite{Jones2009}. The algorithm is deterministic and recursively divides
+the search space and forms a tree of hyper-rectangles (boxes). DIRECT uses the objective function
+values and the Lipschitz condition (\ref{eq:lip}) to estimate promising boxes.
 
-  \item \textbf{Locally-based DIRECT (DIRECT$l$)} \cite{Gablonsky2001}. It's a variation of DIRECT which pays less attention to non-promising boxes and therefore
-  has less exploration power: it can converge faster on problems with few local minimas, but lost the global one in complicated cases.
+  \item \textbf{Locally-based DIRECT (DIRECT$l$)} \cite{Gablonsky2001}. It's a variation of
+DIRECT which pays less attention to non-promising boxes and therefore
+  has less exploration power: it can converge faster on problems with few local minima, but lost the
+global one in complicated cases.
 
-  \item \textbf{Dual Simulated Annealing} \cite{XIANG1997216}.
-This stochastic method is a combination of the Classical Simulated Annealing and the Fast Simulated Annealing coupled to a strategy for applying a local search on accepted locations.
-It converges much faster than both parent algorithms, CSA and FSA.
+  \item \textbf{Dual Simulated Annealing} \cite{XIANG1997216}. This stochastic method is a
+combination of the Classical Simulated Annealing and the Fast Simulated Annealing coupled to a
+strategy for applying a local search on accepted locations. It converges much faster than both parent
+algorithms, CSA and FSA.
 
-  \item \textbf{Differential Evolution} \cite{Storn1997}. DE is an adaptation of the original genetic algorithm to
+  \item \textbf{Differential Evolution} \cite{Storn1997}. DE is an adaptation of the original genetic
+algorithm to
   the continuous search domain.
 
-  \item \textbf{Controlled Random Search} \cite{Price1983}. The CRS starts with a set of random points and then defines
-  the next trial point in relation to a simplex chosen randomly from a stored configuration of points. CRS in not an
-  evolutional algorithm, although stores something like population and performs transformation resembling a mutation.
+  \item \textbf{Controlled Random Search} \cite{Price1983}. The CRS starts with a set of random
+points and then defines
+  the next trial point in relation to a simplex chosen randomly from a stored configuration of points.
+CRS in not an
+  evolutional algorithm, although stores something like population and performs transformation
+resembling a mutation.
 
-  \item \textbf{StoGO} \cite{Madsen1998}. StoGO is dividing the search space into smaller hyper-rectangles via a branch-and-bound approach,
+  \item \textbf{StoGO} \cite{Madsen1998}. StoGO is dividing the search space into smaller hyper-
+rectangles via a branch-and-bound approach,
   and searching them by a local-search algorithm, optionally including some randomness.
@@
 are available in source codes as parts of wide-spread optimization packages.
-AGS, DIRECT, DIRECT$l$, CRS, MLSL and StoGO are part of the NLOpt libray \cite{nlopt}. Differential Evolution and DSA can be found in
+AGS, DIRECT, DIRECT$l$, CRS, MLSL and StoGO are part of the NLOpt library \cite{nlopt}.
+Differential Evolution and DSA can be found in
 the latest version of the SciPy \cite{scipy} package for Python.
@@
 \section{Tools for Comparison of Global Optimization Algorithms}
-\begin{Russian}
 
-Использование сгенерированных некоторыми случайными механизмами
-наборов тестовых задач с известными решениями является одним из общепринятых подходов
-к сравнению алгоритмов оптимизации \cite{Beiranvand2017}. В данной работе
-будем использовать два генератора тестовых задач, порождающих задачи различной природы \cite{grishaginClass, Gaviano2003}.
+The use of the sets of test problems with known solutions generated by some random mechanisms is
+one of commonly accepted approaches to comparing the optimization algorithms
+\cite{Beiranvand2017}. In the present work, we will use two generators of test problems generating
+the problems of different nature \cite{grishaginClass, Gaviano2003}.
 
-Обозначим набор задач, полученных с помощью первого генератора из \cite{grishaginClass} как \(F_{GR}\). Механизм построения задач \(F_{GR}\),
-не предусматривает контроль за сложностью задач и количеством локальных оптимумов, однако известно, что порождаемые функции
-являются существенно многоэкстремальными. Кроме того, задачи, порождаемые \(F_{GR}\) двухмерные. В данной работе будем использовать
-100 случайно сгенерированных функций из класса \(F_{GR}\).
+Let us denote the problem set obtained with the use of the first generator from \cite{grishaginClass}
+as \(F_{GR}\). The mechanism of generation of the problems \(F_{GR}\) doesn't provide the
+control of the problem complexity and of the number of local optima. However, the generated
+functions are known to be the multiextremal ones essentially. Besides, the problems generated by
+\(F_{GR}\) are the two-dimensional ones. In the present work, we will use 100 functions from the
+class \(F_{GR}\) generated randomly.
 
-Генератор GKLS \cite{Gaviano2003} позволяет получать задачи заданной размерности и с заданным количеством экстремумов.
-Кроме того GKLS позволяет регулировать сложность задач, уменьшая или увеличивая размер области притяжения глобального минимума.
-В работе \cite{SergeyevKvasov2006} указаны параметры генератора, позволяющие породить наборы по 100 задач двух уровней сложности
-(Simple и Hard) размерностей 2, 3, 4 и 5. Следуя примеру авторов генератора GKLS, будем использовать предложенные ими параметры и,
-таким образом, добавим в тестовый ещё 800 задач различной размерности и сложности.
+The GKLS generator \cite{Gaviano2003} allows obtaining the problems of given dimensionality
+with given number of extrema. Besides, GKLS allows adjusting the complexity of the problems by
+decreasing or increasing the size of the global minimum attractor. In Ref.
+\cite{SergeyevKvasov2006}, the parameters of the generator allowing generating the sets of 100
+problems each of two levels of complexity (Simple and Hard) of the dimensionality equal to 2, 3, 4,
+and 5 are given. Following the authors of the GKLS generator, we will use the parameters proposed
+by them and, this way, add 800 more problems of various dimensionalities and complexity into the
+test problem set.
 
-Будем считать, что тестовая задача решена, если метод оптимизации провёл очередное испытание \(y^k\) в
-\(\delta\)-окрестности глобального минимума \(y^*\), т.е. $\left\|y^k-
-y^*\right\|\leqslant \delta = 0.01\left\|b-a\right\|$, где \(a\) и \(b\) --- левая и правая границы гиперкуба из (\ref{eq:task}).
-Если указанное соотношение не выполнено до истечения лимита на количество испытаний, то задача считается нерешённой.
-Максимальный лимит на количество испытаний установлен для каждого класса задач, в соответствии с размерностью и сложностью (см. таблицу \ref{tab:limits}).
+Let us suppose a test problem to be solved if the optimization method executes the scheduled trial
+\(y^k\) in a \(\delta\)-nearness of the global minimum \(y^*\) i. e. $\left\|y^k-y^*\right\|\leqslant \delta
+= 0.01\left\|b-a\right\|$ where \(a\) and \(b\) are the left and the right boundaries of the hypercube
+from (\ref{eq:task}). If this relation is not fulfilled before the expiration of the limit of the number of
+trials, the problem was considered to be unsolved. The maximum limit of the number of trials was set
+for each problem class according to the dimensionality and complexity (see Table \ref{tab:limits}).
 
@@
 \begin{center}
-\caption{Trials limit for the each of test problems classes}
+\caption{Trials limits for the test problem classes}
   \begin{tabular}{|l|{c}|}
@@
 
-В качестве характеристик метода оптимизации на каждом из классов будем рассматривать среднее число
-испытаний, затраченное для решения одной задачи, и количество решённых задач. Чем меньше число испытаний, тем быстрее метод сходится
-к решению, а значит и меньше обращается к потенциально трудоёмкой процедуре вычислений целевой функции.
-Количество решённых задач говорит о надёжности метода с заданными параметрами на решаемом классе тестовых задач.
-Чтобы сделать величины, характеризующие надёжность и скорость сходимости, независимыми при подсчёте среднего количества испытаний учитывались только решённые задачи.
-\end{Russian}
+Let us consider the averaged number of trials executed to solve a single problem and the number of
+solved problems as the characteristics of the optimization method on each class. The less the number
+of trials, the faster the method converges to a solution, hence the less times it turns to a potentially
+computation-costly procedure of computing the objective function. The number of solved problems
+evidences the reliability of the method at given parameters on the class of test problems being
+solved. In order to calculate the quantities featuring the reliability and the speed of convergence,
+when calculating the averaged number of trials the solved problems were taken into account only.
 
@@
 \subsection{Results of Sequential Algorithms}
-\begin{Russian}
 
-Результаты тех или иных алгоритмов на различных классах задач напрямую зависят от
-настроек алгоритмов. В большинстве случаев авторы программных реализаций ориентируются на
-задачи средней сложности и чтобы получить удовлетворительный результат при решении
-существенно многоэкстремальных задач, требуется корректировка некоторых параметров.
-При проведении сравнения были заданы следующие параметры для методов:
+The results of various algorithms on different problem classes depend on the adjustments of
+algorithms directly. In most cases, the authors of software implementations are oriented onto the
+problems of medium difficulty. In order to obtain a satisfactory result when solving the essentially
+multiextremal problems, a correction of some parameters is required. When conducting the
+comparison, the following parameters for the methods were employed:
 \begin{itemize}
-  \item в методе AGS\(l\) параметр чередования глобального и локального правил вычисления характеристик был задан равным 5:1;
-  \item в методах DIRECT и DIRECT\(l\) параметр \(\epsilon=10^{-4}\);
-  \item в методе SDA параметр \(visit=2.72\).
+  \item in the AGS\(l\) method, the parameter of alternation of the global rules of computing the
+characteristic and the local one was set to be equal to 5:1;
+  \item in the DIRECT and DIRECT\(l\) methods, the parameter \(\epsilon=10^{-4}\);
+  \item in the SDA method, the parameter \(visit=2.72\).
 \end{itemize}
 
-Остальные параметры варьировались в зависимости от класса задач (см. таблицу \ref{tab:params}).
+The rest parameters were varied subject to the problem class (see Table \ref{tab:params}).
 
@@
 
-В таблицах \ref{tab:trials}, \ref{tab:solved} приведены результаты запуска методов оптимизации на рассматриваемых классах задач.
-Методы DIRECT и AGS\(l\) показали наилучшую скорость сходимости на всех классах, причём AGS\(l\) уступает
-DIRECT на задачах из классов Simple и обходит его на задачах классов Hard. Как видно из таблицы \ref{tab:solved},
-самыми надёжными методами являются детерминированные: AGS, AGS\(l\), DIRECT и DIRECT\(l\). Из стохастических методов
-самую высокую надёжность демонстрируют MLSL и SDA.
+The results of running the optimization methods on the considered problem classes are presented in
+Tables \ref{tab:trials}, \ref{tab:solved}. The DIRECT and AGS\(l\) methods have demonstrated the
+best convergence speed on all classes, at that AGS\(l\) inferior to DIRECT on the problems from the
+Simple classes and has an advantage on the problems of the Hard classes. As one can see from Table
+\ref{tab:solved}, the deterministic methods (AGS, AGS\(l\), DIRECT, and DIRECT\(l\)) were the
+most reliable. Among the stochastic methods, MLSL and SDA have demonstrated the highest
+reliability.
 
@@
   \hline
-  \(F_{GR}\)     & 193.1 &  \textbf{158.3} & 400.3 & 182.3 & 214.9 & 947.2 & 691.2 & 1257.3 & 1336.8 \\
+  \(F_{GR}\)     & 193.1 &  \textbf{158.3} & 400.3 & 182.3 & 214.9 & 947.2 & 691.2 & 1257.3 &
+1336.8 \\
   \hline
-  GKLS 2d Simple &  254.9 & 217.6 & 510.6 & \textbf{189.0} & 255.2 & 556.8 & 356.3 & 952.2 & 1251.5 \\
+  GKLS 2d Simple &  254.9 & 217.6 & 510.6 & \textbf{189.0} & 255.2 & 556.8 & 356.3 & 952.2
+& 1251.5 \\
   \hline
-  GKLS 2d Hard   &  728.7 & \textbf{488.0} & 844.7 & 985.4 & 1126.7 & 1042.5 & 1637.9 & 1041.1 & 2532.2 \\
+  GKLS 2d Hard   &  728.7 & \textbf{488.0} & 844.7 & 985.4 & 1126.7 & 1042.5 & 1637.9 &
+1041.1 & 2532.2 \\
   \hline
-  GKLS 3d Simple &  1372.1 & 1195.3 & 4145.8 & \textbf{973.6} & 1477.8 & 4609.2 & 2706.5 & 5956.94 & 3856.1 \\
+  GKLS 3d Simple &  1372.1 & 1195.3 & 4145.8 & \textbf{973.6} & 1477.8 & 4609.2 & 2706.5 &
+5956.94 & 3856.1 \\
   \hline
-  GKLS 3d Hard   &  3636.1 & \textbf{1930.5} & 6787.0 & 2298.7 & 3553.3 & 5640.1 & 4708.4 & 6914.3 & 7843.2 \\
+  GKLS 3d Hard   &  3636.1 & \textbf{1930.5} & 6787.0 & 2298.7 & 3553.3 & 5640.1 & 4708.4 &
+6914.3 & 7843.2 \\
   \hline
-  GKLS 4d Simple &  26654.1 & 11095.7 & 37436.8 & \textbf{7824.3} & 15994.1 & 41514.3 & 21417.9 & 19157.7 & 59895.4 \\
+  GKLS 4d Simple &  26654.1 & 11095.7 & 37436.8 & \textbf{7824.3} & 15994.1 & 41514.3 &
+21417.9 & 19157.7 & 59895.4 \\
   \hline
-  GKLS 4d Hard   &  54536.8 &  \textbf{23167.8} &  73779.3 &  23204.4 &  54489.9 &  80247.2 &  68815.5 &  27466.1 &  109328.1  \\
+  GKLS 4d Hard   &  54536.8 &  \textbf{23167.8} &  73779.3 &  23204.4 &  54489.9 &  80247.2
+&  68815.5 &  27466.1 &  109328.1  \\
   \hline
-  GKLS 5d Simple &  29810.0 & 11529.0 & 143575.0 & \textbf{7166.5} & 13970.5 & 52647.6 & 34255.3 & 73074.5 & 91580.4 \\
+  GKLS 5d Simple &  29810.0 & 11529.0 & 143575.0 & \textbf{7166.5} & 13970.5 & 52647.6 &
+34255.3 & 73074.5 & 91580.4 \\
   \hline
-  GKLS 5d Hard   &  113129.1 & 67652.7 & 165192.8 & \textbf{66327.4} & 164390.6 & 138766.2 & 116973.1 & 105496.9 & 155123.8 \\
+  GKLS 5d Hard   &  113129.1 & 67652.7 & 165192.8 & \textbf{66327.4} & 164390.6 & 138766.2
+& 116973.1 & 105496.9 & 155123.8 \\
   \hline
@@
 
-The computational experiments with parallel algorithms have been carried out on the Lobachevsky supercomputer at
+The computational experiments with parallel algorithms have been carried out on the Lobachevskii
+supercomputer at
 State University of Nizhni Novgorod. A computational node includes 2 Intel
@@
 
-Сравнение двух параллельных алгоритмов, реализованных в системах Globalizer и MIDACO
-было проведено на четырёхмерных задачах GKLS, т.к. задачи большей размерности не поддерживаются
-бесплатной версией MIDACO. Как и в случае с последовательными алгоритмами, для кадого класса задач
-были изменены параметры методов. Для MIDACO параметр \(focus = -1\). В случае с Globalizer параметр
-\(r\) был выбран следуя таблице \ref{tab:params}. С удвоением количества задействованных развёрток
-этот параметр уменьшался на 0.3, поскольку увеличение количества развёрток увеличивает и надёжность метода из секции \ref{sub:parallel_evolvents}
-(подробнее см. \cite{SOVRASOV2018}).
+The comparison of two parallel algorithms implemented in the Globalizer and MIDACO systems
+was conducted on four-dimensional GKLS problems since the problems of larger dimensionality are
+not supported by the free version of MIDACO. As in the case of the sequential algorithms, the
+parameters of methods were determined for each problem class. For MIDACO, the parameter
+\(focus = -1\). In the case of Globalizer, the parameter \(r\) was selected following Table
+\ref{tab:params}. This parameter decreased by 0.3 with doubling the number of employed evolvents
+since the increasing of the number of evolvents enhances the reliability of the method from Section
+\ref{sub:parallel_evolvents} as well (see \cite{SOVRASOV2018} for details).
 
-Если считать, что затраты на параллелизм пренебрежимо малы по сравнению с затратами на вычисление целевых функций в задачах оптимизации, то ускорение по времени от использования параллельного метода будет равно ускорению по итерациям.
-Однако, в действительности это предположение справедливо,
-тлолько если вычисление значения целевой функции достаточно трудоёмко. Во всех численных экспериментах время вычисления целевой функции занимает примерно $2\times 10^{-3}$с.
+If one considers the costs of parallelism to be negligible as compared to the costs of computing the
+objective functions in the optimization problems, the speedup in time due to the use of the parallel
+method would be equal to the speedup with respect to the number of iterations. However, actually
+this assumption is true if the computing of the objective function is well computation costly only. In
+all numerical experiments, the time of computing the objective function was approximately $2.5\times
+10^{-3}$ s.
 
@@
 The number of iterations is reduced considerably with increasing the number of nodes and the
-number of threads on each node. В скобках также указано количество решённых задач в каждом классе.
-Globalizer решает практически все задачи в обоих классах, в то время как MIDACO не решает многие задачи из класса GKLS 4d Hard.
-Стоит ещё отметить, что с ростом числа узлов надёжность MIDACO падает на обоих классах задач, в то время как надёжность Globalizer остаётся стабильной.
+number of threads on each node. Also, the number of solved problems in each class is given  in
+brackets.
+Globalizer solves almost all problems in both classes whereas MIDACO hasn't solved a considerable
+number of problems from the GKLS 4d Hard class. Also, it is worth noting that the reliability of
+MIDACO decreased with increasing number of nodes on both problem classes whereas the
+reliability of Globalizer remains stable.
 
@@
 
-Среднее ускорение по итерациям и среднее ускорение по времени (в скобках) приведены в таблице \ref{tab:speedup}.
-В первой строке таблицы приведено среднее количество итераций и среднее время решения задачи в
-последовательном режиме (в скобках). MIDACO для при параллельных вычислениях использует
-меньшее количество коммуникаций, поэтому показатели ускорения монотонно возрастают с ростом количества
-вычислительных узлов. В Globalizer каждый узел должен переслать результаты проведения итерации
-всем другим узлам и с увеличением количества узлов и количества рабочих потоков на каждом узле
-эффективность распараллеливания падает. Это можно видеть в таблице \ref{tab:speedup}, когда при переходе от 4х
-узлов к 8 ускорение по времени при использовании 16 потоков упало. В случае одного потока на узел
-ускорение обоих методов оптимизации примерно одинаковое. Однако стоит заметить, что Globalizer выполняет
-меньше испытаний и при \(p=1\) тратит гораздо меньше времени на решение всех задач.
-С ростом количества узлов и потоков MIDACO догоняет Globalizer по среднему времени решения задачи,
-однако надёжность при этом падает.
+The averaged speedup in the number of iterations and the one in time (in brackets) are presented in
+Table \ref{tab:speedup}. In the first row of Table \ref{tab:speedup}, the averaged number of
+iterations and the averaged time of solving the problems in the sequential mode (in brackets) are
+presented. MIDACO uses a smaller number of communications in the parallel computations.
+Therefore, the indicators of speedup increased monotonously with increasing number of
+computational nodes employed. In Globalizer, each node should transfer the results of the executed
+iterations to all other nodes. Therefore, the efficiency of parallelization decreased with increasing
+number of nodes and the number of active threads on each node. This is evident from Table
+\ref{tab:speedup}: the speedup in time decreased at the transition from 4 nodes to 8 ones when
+using 16 threads. In the case of using 1 thread per a node, the speedup of both optimization methods
+were approximately the same. However, it should be noted that Globalizer executed less number of
+trials, and spends much less time for solving all problems when \(p=1\). With increasing number of
+nodes and threads, MIDACO approaches Globalizer in the averaged time of problem solving,
+however, its reliability decreases at the same time.
 
-Аномальное сверхлинейное ускорение по времени
-MIDACO на классе GKLS 2d Hard при использовании 8 и 12 узлов можно объяснить тем, что, согласно таблице
-\ref{tab:iterations}, MIDACO решил больше задач в параллельном режиме, чем в последовательном, а значит, на
-меньшем количестве задач работал до исчерпания лимита испытаний.
+An anomalous superlinear speedup of MIDACO in time on the GKLS 2d Hard class when
+employing 8 and 12 nodes can be explained by the fact that, according to Table \ref{tab:iterations},
+MIDACO has solved a greater number of problems in the parallel mode than in the sequential one,
+hence, has worked until the limit of trials is expired on a less number of problems.
 
@@
 
-В статье были рассмотрены несколько последовательных и параллельных алгоритмов глобальной оптимизации.
-Было проведено сравнение эффективности их работы на множестве тестовых задач, по результатам которого можно сделать
-следующие выводы:
+In the present paper, several sequential and parallel global optimization algorithms were considered.
+A comparison of efficiencies of these ones has been done on a set of test problems. The results allow
+making the following conclusions:
 \begin{itemize}
-  \item методы AGS\(l\), реализованный в системе Globalizer показал скорость сходимости и надёжность на уровне \(DIRECT\) и
-  превосходит многие другие алгоритмы, реализации которых есть в открытом доступе;
-  \item стохастические методы оптимизации проигрывают детерминированным по скорости сходимости и надёжности. Особенно
-  сильно это проявляется на более сложных многоэкстремальных задачах;
-  \item параллельная версия системы Globalizer демонстрирует хорошие показатели ускорения по времени при работе на нескольких узлах,
-  на каждом из которых вычисляется одно значение целевой за итерацию. При решении задач с быстро вычислимой целевой функцией и
-  использовании многопоточности на узлах показатели ускорения системы Globalizer деградируют с увеличением количества узлов;
-  \item система MIDACO больше всего подходит для несложных глобальных задач, целевые функции в которых могут быть быстро вычислимы.
-  В таком случае MIDACO достаточно надёжна и обеспечивает линейное ускорение с ростом количества узлов и параллельно выполняемых потоков на них.
+  \item AGS\(l\) method implemented in the Globalizer system has demonstrated the convergence
+speed and reliability at the level of \(DIRECT\) and exceeds many other algorithms, the open-access
+implementations of which are available;
+  \item the stochastic optimization methods inferior to the deterministic ones in the convergence
+speed and in reliability. It is manifested especially strongly on more complex multiextremal
+problems;
+  \item the parallel version of the Globalizer system demonstrates good indicators of speedup in time
+when running on several nodes, on each of which a single objective function value per iterations is
+computed. When solving the problems with fast computable objective functions and using multiple
+threads on the nodes, the indicators of speedup for the Globalizer system degrade with increasing
+number of nodes;
+  \item the MIDACO system is the most suitable for simple global optimization problems with the
+fast-computable objective functions. In this case, MIDACO is reliable enough and provides a linear
+speedup with increasing number of nodes and threads executed on these ones in parallel.
 \end{itemize}
-
-\end{Russian}
 

### 2018-11-28 2019/PAVT/pavt_paper.tex
PAVT: add figures

@@
 \usepackage{cite}
+\usepackage{subfig}
 % to typeset URLs, URIs, and DOIs
@@
 algorithms. In the first part of the paper popular sequential open-source solvers are compared against
-the Globalizer solver, which is developed at the Lobachevskii State University. The Globalizer is
+the Globalizer solver, which is developed at the Lobachevsky State University. The Globalizer is
 designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows
@@
 evidences the reliability of the method at given parameters on the class of test problems being
-solved. In order to calculate the quantities featuring the reliability and the speed of convergence,
-when calculating the averaged number of trials the solved problems were taken into account only.
+solved. In order to make independent the quantities featuring the reliability and the speed of convergence,
+averaged number of trials always was calculated taking into account solved problems only.
 
@@
 
-The computational experiments with parallel algorithms have been carried out on the Lobachevskii
+Computational experiments with parallel algorithms have been carried out on the Lobachevsky
 supercomputer at
@@
 Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. Each CPU has 8 cores (i. e. total 16
-cores were available per a node). All computational nodes are working under CentOS Linux 7.2.
-All considered algorithms are implemented using C++ and compiled by GCC 4.8.5.
+cores were available per a node). All the computational nodes are working under CentOS Linux 7.2.
+All the considered algorithms are implemented using C++ and compiled by GCC 4.8.5.
 
@@
   \centering
-  \caption{Averaged numbers of iterations executed by the parallel algorithm for solving the test
+  \caption{Averaged numbers of iterations executed by the parallel algorithms for solving the test
 optimization problems}
@@
 number of nodes and the number of active threads on each node. This is evident from Table
-\ref{tab:speedup}: the speedup in time decreased at the transition from 4 nodes to 8 ones when
+\ref{tab:speedup} and Picture \ref{fig:speedup_p16}: the speedup in time decreased at the transition from 4 nodes to 8 ones when
 using 16 threads. In the case of using 1 thread per a node, the speedup of both optimization methods
-were approximately the same. However, it should be noted that Globalizer executed less number of
+were approximately the same (see Picture \ref{fig:speedup_p16}). However, it should be noted that Globalizer executed less number of
 trials, and spends much less time for solving all problems when \(p=1\). With increasing number of
-nodes and threads, MIDACO approaches Globalizer in the averaged time of problem solving,
+nodes and threads, MIDACO approaches Globalizer in the averaged time of problem solving (see Pictures \ref{fig:time_p1}, \ref{fig:time_p16}),
 however, its reliability decreases at the same time.
@@
 
+\begin{figure}[ht]
+  \centering
+  \subfloat[$p=1$]{{\includegraphics[width=.5\textwidth]{images/speedup_midaco_globalizer_p_1.pdf}}\label{fig:speedup_p1}}
+  \subfloat[$p=16$]{{\includegraphics[width=.5\textwidth]{images/speedup_midaco_globalizer_p_16.pdf}}\label{fig:speedup_p16}}
+  \caption{Speedup demonstrated by the parallel algorithms when solving problems from the GKLS 4d Hard class}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \subfloat[$p=1$]{{\includegraphics[width=.5\textwidth]{images/time_midaco_globalizer_p_1.pdf}}\label{fig:time_p1}}
+  \subfloat[$p=16$]{{\includegraphics[width=.5\textwidth]{images/time_midaco_globalizer_p_16.pdf}}\label{fig:time_p16}}
+  \caption{Averaged execution time of the parallel algorithms when solving problems from the GKLS 4d Hard class}
+\end{figure}
+
 \begin{table}
   \centering
-  \caption{Speedup of parallel computations executed by the parallel algorithm}
+  \caption{Speedup of parallel computations executed by the parallel algorithms}
   \label{tab:speedup}

### 2018-11-29 2019/PAVT/pavt_paper.tex
PAVT: update

@@
 algorithms. In the first part of the paper popular sequential open-source solvers are compared against
-the Globalizer solver, which is developed at the Lobachevsky State University. The Globalizer is
+Globalizer solver, which is developed at the Lobachevsky State University. The Globalizer is
 designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows
@@
 
-The nonlinear global optimization of the non-convex functions is considered to be one of the most
-difficult problems of mathematical programming traditionally. The finding of the global minimum of
-a function of several variables appears to be more difficult than the local optimization in the
-thousand-dimensional space often. For the latter, the application of the simplest gradient descent
-method may appear to be enough whereas in order to \textit{guarantee} the finding of the global
-optimum by the optimization methods, one has to accumulate the information on the behavior of the
+Nonlinear global optimization of non-convex functions is traditionally considered to be one of the most
+difficult problems of mathematical programming. The finding of the global minimum of
+a function of several variables often appears to be more difficult than local optimization in the
+thousand-dimensional space. In local optimization applying of the simplest gradient descent
+method may appear to be enough, whereas in order to \textit{guarantee} find the global
+optimum, an optimization method has to accumulate information about behavior of the
 objective function within the whole search domain
-\cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}. To date, various stochastic global
+\cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}. Nowadays, various stochastic global
 optimization algorithms have attracted much attention, first of all, the evolution ones
-\cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}. These ones have rather simple structure
+\cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}. They have rather simple structure
 and allow solving the problems of large dimensionality. However, the evolution algorithms ensure
@@
 
-In the present paper, the open source implementations of nine different global optimization methods
-were considered. All algorithms were tested using a set of 900 essentially multiextremal functions,
-which has been generated using special problem generators \cite{Gaviano2003, grishaginClass}.
+In the present paper, the open-source implementations of nine different global optimization methods
+were considered. All the algorithms were tested using a set of 900 essentially multiextremal functions,
+which has been generated using generators of test problems \cite{Gaviano2003, grishaginClass}.
 Besides the comparison of the sequential algorithms, a comparison of the MIDACO solver
-\cite{Schlueter2012} with the Globalizer software system \cite{globalizerSystem,Strongin2018} has been
-performed while running on Lobachevsky supercomputer using a subset of 200 test problems.
+\cite{Schlueter2012} with the Globalizer software system \cite{globalizerSystem,Strongin2018} on a subset of 200 test problems has been
+performed in the Lobachevsky supercomputer environment.
 
@@
 as well as of the deterministic ones \cite{posik2012, KVASOV2018245, Liberti2005} between each
-other has been considered in the literature. In these works, the majority of modern methods have
+other has been considered in the literature. In these works, most of modern methods have
 been studied in details, but the emphasis was made mainly on the sequential algorithms while the
@@
 methods. In the majority of works, the sets of well-known test problems (for example, the Rastrigin
-functions, Ackley function, etc.) were taken as the sets of test functions. The sizes of such sets don't
+function, Ackley function, etc.) were taken as the sets of test functions. The sizes of such sets don't
 exceed 100 different functions usually, some of which can be the single-extremal ones (such as the
-Rosenbrock functions).
+Rosenbrock function).
 
@@
 \[
-\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
+\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leq s\leq S
 \]
@@
 \par
-3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and
+3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leq s\leq S\), of the point \(y^k\) and
 interpret the
@@
 Let us suppose a test problem to be solved if the optimization method executes the scheduled trial
-\(y^k\) in a \(\delta\)-nearness of the global minimum \(y^*\) i. e. $\left\|y^k-y^*\right\|\leqslant \delta
+\(y^k\) in a \(\delta\)-nearness of the global minimum \(y^*\) i. e. $\left\|y^k-y^*\right\|\leq \delta
 = 0.01\left\|b-a\right\|$ where \(a\) and \(b\) are the left and the right boundaries of the hypercube
@@
 \begin{center}
-\caption{Class-specific parameters of optimization algorithms}
+\caption{Class-specific parameters of the optimization algorithms}
   \begin{tabular}{|l|{c}|{c}|{c}|}
@@
 Computational experiments with parallel algorithms have been carried out on the Lobachevsky
-supercomputer at
-State University of Nizhni Novgorod. A computational node includes 2 Intel
+supercomputer at State University of Nizhni Novgorod. A computational node includes 2 Intel
 Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. Each CPU has 8 cores (i. e. total 16
 cores were available per a node). All the computational nodes are working under CentOS Linux 7.2.
-All the considered algorithms are implemented using C++ and compiled by GCC 4.8.5.
+All the considered algorithms are implemented using C++ and compiled by the GCC 4.8.5.
 
@@
 iterations and the averaged time of solving the problems in the sequential mode (in brackets) are
-presented. MIDACO uses a smaller number of communications in the parallel computations.
+presented. MIDACO uses a smaller number of communications in the parallel computations than the Globalizer.
 Therefore, the indicators of speedup increased monotonously with increasing number of
@@
 \ref{tab:speedup} and Picture \ref{fig:speedup_p16}: the speedup in time decreased at the transition from 4 nodes to 8 ones when
-using 16 threads. In the case of using 1 thread per a node, the speedup of both optimization methods
+using 16 threads. In the case of using 1 thread per a node, the speedups of both optimization methods
 were approximately the same (see Picture \ref{fig:speedup_p16}). However, it should be noted that Globalizer executed less number of

### 2018-11-30 2019/PAVT/pavt_paper.tex
Some corrections. not final

@@
 \begin{abstract}
-This work considers several stochastic and deterministic derivative-free global optimization
+This work considers several stochastic and deterministic deri-vative-free global optimization
 algorithms. In the first part of the paper popular sequential open-source solvers are compared against
@@
 memory. MIDACO is a state-of-the-art global solver included to the TOMLAB optimization
-environment for MATLAB. Results of the benchmark show advantages of the Globalizer on small-
-dimensional, but sufficiently multi-extremal benchmark functions.
+environment for MATLAB. Results of the benchmark show advantages of the Globalizer on small-dimensional, but sufficiently multi-extremal benchmark functions.
 % We would like to encourage you to list your keywords within
@@
 
-Nonlinear global optimization of non-convex functions is traditionally considered to be one of the most
+Nonlinear global optimization of non-convex functions is considered to be one of the most
 difficult problems of mathematical programming. The finding of the global minimum of
-a function of several variables often appears to be more difficult than local optimization in the
+a function of several variables often appears to be more difficult than local optimization in a
 thousand-dimensional space. In local optimization applying of the simplest gradient descent
-method may appear to be enough, whereas in order to \textit{guarantee} find the global
+method may appear to be enough, whereas in order to \textit{guarantee} finding the global
 optimum, an optimization method has to accumulate information about behavior of the
@@
 \subsection{Parallel Algorithm of Global Search}
-\subsubsection{Dimension Reduction with Evolvents}
+\subsubsection{Dimension Reduction with Space-Filling Curves}
 Within the framework of the information-statistical global optimization theory,
-the Peano space-filling curves (or evolvents) \(y(x)\) mapping the interval \([0,1]\)
+the Peano space-filling curves (or \textit{evolvents}) \(y(x)\) mapping the interval \([0,1]\)
 onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
@@
 onto the corresponding one-dimensional intervals. This method for building a set of mappings can be
-''scaled'' easily to obtain more evolvents (up to
+``scaled'' easily to obtain more evolvents (up to
 $2^N$) if necessary.
@@
 \[
-\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leq s\leq S
+\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leq s\leq S,
 \]
@@
 
-  \item \textbf{Locally-based Algorithm of Global Search (AGS\(l\))} \cite{indexMethod}. It's a
+  \item \textbf{Locally-biased Algorithm of Global Search (AGS\(l\))} \cite{indexMethod}. It's a
 modification of
@@
 
-  \item \textbf{Locally-based DIRECT (DIRECT$l$)} \cite{Gablonsky2001}. It's a variation of
+  \item \textbf{Locally-biased DIRECT (DIRECT$l$)} \cite{Gablonsky2001}. It's a variation of
 DIRECT which pays less attention to non-promising boxes and therefore
@@
 
-All the mentioned algorithms (except of AGS\(l\) which implemented in the Globalizer system)
-are available in source codes as parts of wide-spread optimization packages.
-AGS, DIRECT, DIRECT$l$, CRS, MLSL and StoGO are part of the NLOpt library \cite{nlopt}.
+All the mentioned algorithms are available in source codes as parts of wide-spread optimization packages (except of AGS and AGS\(l\), which are implemented in the Globalizer system).
+DIRECT, DIRECT$l$, CRS, MLSL and StoGO are part of the NLOpt library \cite{nlopt}.
 Differential Evolution and DSA can be found in

### 2018-12-01 2019/PAVT/pavt_paper.tex
Some minor corrections

@@
 The GKLS generator \cite{Gaviano2003} allows obtaining the problems of given dimensionality
-with given number of extrema. Besides, GKLS allows adjusting the complexity of the problems by
-decreasing or increasing the size of the global minimum attractor. In Ref.
-\cite{SergeyevKvasov2006}, the parameters of the generator allowing generating the sets of 100
+with given number of extrema. Moreover, GKLS allows adjusting the complexity of the problems by
+decreasing or increasing the size of the global minimum attractor. In 
+\cite{SergeyevKvasov2006} the parameters of the generator allowing generating the sets of 100
 problems each of two levels of complexity (Simple and Hard) of the dimensionality equal to 2, 3, 4,
@@
 Let us suppose a test problem to be solved if the optimization method executes the scheduled trial
-\(y^k\) in a \(\delta\)-nearness of the global minimum \(y^*\) i. e. $\left\|y^k-y^*\right\|\leq \delta
-= 0.01\left\|b-a\right\|$ where \(a\) and \(b\) are the left and the right boundaries of the hypercube
+\(y^k\) in a \(\delta\)-vicinity of the global minimum \(y^*\), i.e. $\left\|y^k-y^*\right\|\leq \delta
+= 0.01\left\|b-a\right\|$, where \(a\) and \(b\) are the left and the right boundaries of the hypercube
 from (\ref{eq:task}). If this relation is not fulfilled before the expiration of the limit of the number of
-trials, the problem was considered to be unsolved. The maximum limit of the number of trials was set
+trials, the problem was considered to be unsolved. The limit of the number of trials was set
 for each problem class according to the dimensionality and complexity (see Table \ref{tab:limits}).
@@
 \section{Results of Numerical Experiments}
-\subsection{Results of Sequential Algorithms}
+\subsection{Sequential Algorithms}
 
@@
 \begin{itemize}
-  \item in the AGS\(l\) method, the parameter of alternation of the global rules of computing the
-characteristic and the local one was set to be equal to 5:1;
+  \item in the AGS\(l\) method, the parameter of alternation the
+global and local iterations was set to be equal to 5:1;
   \item in the DIRECT and DIRECT\(l\) methods, the parameter \(\epsilon=10^{-4}\);
@@
 
-\subsection{Results of Parallel Algorithms}
+\subsection{Parallel Algorithms}
 
@@
   \item AGS\(l\) method implemented in the Globalizer system has demonstrated the convergence
-speed and reliability at the level of \(DIRECT\) and exceeds many other algorithms, the open-access
+speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-access
 implementations of which are available;
@@
 problems;
-  \item the parallel version of the Globalizer system demonstrates good indicators of speedup in time
+  \item the parallel version of the Globalizer system demonstrates good speedup
 when running on several nodes, on each of which a single objective function value per iterations is
 computed. When solving the problems with fast computable objective functions and using multiple
-threads on the nodes, the indicators of speedup for the Globalizer system degrade with increasing
+threads on the nodes, the speedup for the Globalizer system degrade with increasing
 number of nodes;
@@
 \section*{Acknowledgements}
-The study was supported by the Russian Science Foundation, project No 16-11-
-10150.
+The study was supported by the Russian Science Foundation, project No 16-11-10150.
 

### 2018-12-03 2019/PAVT/pavt_paper.tex
PAVT: update pdf

@@
   \item AGS\(l\) method implemented in the Globalizer system has demonstrated the convergence
-speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-access
+speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-source
 implementations of which are available;

### 2018-12-05 2018/Journal of Supercomputing/JoS_Paper.tex
References are corrected

@@
 
-%\bibitem{Balandin2011} Balandin, D.V., Kogan, M.M.: Optimal linear-quadratic control: 
-From matrix equations to linear matrix inequalities. Automation and Remote Control. 72(11), 
-2276--2284 (2011)
+%\bibitem{Balandin2011} Balandin, D.V., Kogan, M.M.: Optimal linear-quadratic control: From matrix equations to linear matrix inequalities. Automation and Remote Control. 72(11), 2276--2284 (2011)
 
-%\bibitem{Balandin2017}Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-
-control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 
-(2017)
+%\bibitem{Balandin2017}Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 (2017)
 
@@
 
-%\bibitem{Barkalov2015} Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi coprocessor 
-for solving global optimization problems. LNCS. 9251, 307--318 (2015)
+%\bibitem{Barkalov2015} Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi coprocessor for solving global optimization problems. LNCS. 9251, 307--318 (2015)
 
@@
 \bibitem{Barkalov2017}
-Barkalov, K., Lebeled, I.: Comparing Two Approaches for Solving Constrained Global 
+Barkalov, K., Lebelev, I.: Comparing Two Approaches for Solving Constrained Global 
 Optimization Problems. LNCS. 10556, 301--306 (2017)

### 2018-12-11 2019/ICCS_Barkalov/ICCS_Paper.tex

--- 2018-12-05 2018/Journal of Supercomputing/JoS_Paper.tex
@@
 %
-\RequirePackage{fix-cm}
+
 %
@@
 %\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
-\documentclass[smallextended]{svjour3}       % onecolumn (second format)
+%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
 %\documentclass[twocolumn]{svjour3}          % twocolumn
+\documentclass{llncs}
 
@@
 
-\smartqed  % flush right qed marks, e.g. at end of proof
+% flush right qed marks, e.g. at end of proof
 %
@@
 \usepackage{multirow}
+\usepackage{hyperref}
+\usepackage[misc,geometry]{ifsym}
+
+\usepackage{amsmath}
+
 %
@@
 % Insert the name of "your journal" with
- \journalname{The Journal of Supercomputing}
+
 %
@@
 
-\title{Parallel Computing for Constrained Global Optimization Problems
-\thanks{The study was supported by the Russian Science Foundation, project No 16-11-
-10150.}
-}
+\mainmatter
+
+\title{Parallel Computing for Constrained Global Optimization Problems}
+
 %\subtitle{Do you have a subtitle?\\ If so, write it here}
@@
 
-\author{Konstantin Barkalov          \and
-        Ilya Lebedev %etc.
+\author{Konstantin Barkalov \Letter \and
+        Ilya Lebedev %\Letter
+\\
+\email{konstantin.barkalov@itmm.unn.ru}
+%\\
+%\email{ilya.lebedev@itmm.unn.ru}
 }
@@
 
-\institute{K. Barkalov  \at
-               Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
-              \email{konstantin.barkalov@itmm.unn.ru} 
-           \and
-           I. Lebedev \at
-					Lobachevsky State University of Nizhni Novgorod, Nizhni 
-Novgorod, Russia \\
-							\email{ilya.lebedev@itmm.unn.ru} 
+\institute{
+Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
 }
-
-\date{Received: date / Accepted: date}
-% The correct dates will be entered by the editor
 
@@
 constrains generated randomly.
-\keywords{Global optimization \and Constrained problems \and Non-convex constraints \and 
-Dimensionality reduction \and Parallel algorithms}
+
+\keywords global optimization, constrained problems, non-convex constraints, 
+dimensionality reduction, parallel algorithms.
 % \PACS{PACS code1 \and PACS code2 \and more}
 % \subclass{MSC code1 \and MSC code2 \and more}
+
 \end{abstract}
 
-\section{Introduction}
-\label{intro}
+\section{Introduction}\label{intro}
 
@@
 
-%Another example is the optimal control problem described through system of ordinary 
-%differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}. It is only 
-%possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz 
-%matrix, i.e. every eigenvalue of $A$ has a strictly negative real part. Otherwise, the value of the 
-%criteria is indeterminate.
+Another example is the optimal control problem described through system of ordinary 
+differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}. It is only 
+possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz 
+matrix, i.e. every eigenvalue of $A$ has a strictly negative real part. Otherwise, the value of the 
+criteria is indeterminate.
 
@@
 
+\section*{Acknowledgments}
+The study was supported by the Russian Science Foundation, project No 16-11-10150.
+
 % Non-BibTeX users please use
-\begin{thebibliography}{}
+\begin{thebibliography}{99}
 
@@
 Ferreiro, A.M., Garcia, J.A., Lopez-Salas, J.G., Vazquez, C.: An efficient implementation of 
-parallel simulated annealing algorithm in GPUs. J. Glob. Optim. 57(3), 863--890 (2013)
+parallel simulated annealing algorithm in GPUs. J. Glob. Optim. \textbf{57}(3), 863--890 (2013)
 
@@
 hardware acceleration: an investigation on bound constrained optimization problems. J. Glob. 
-Optim. 50(3), 417--437 (2011).
+Optim. \textbf{50}(3), 417--437 (2011).
 
@@
 Koro\v sec, P., Vajter\v sic, M.,  \v Silc, J., Kutil, R.: Multi-core implementation of the 
-differential ant-stigmergy algorithm for numerical optimization. J. Supercomput. 63(3), 757--
+differential ant-stigmergy algorithm for numerical optimization. J. Supercomput. \textbf{63}(3), 757--
 772 (2013)
@@
 Guerrero, G.D., Cecilia, J.M., Llanes, A. et al.: Comparative evaluation of platforms for parallel 
-Ant Colony Optimization. J. Supercomput. 69(1) 318--329 (2014)
+Ant Colony Optimization. J. Supercomput. \textbf{69}(1) 318--329 (2014)
 
@@
 Paulavi\v cius, R., \v Zilinskas, J., Grothey, A.: Parallel branch and bound for global 
-optimization with combination of Lipschitz bounds. Optim. Methods Softw. 26(3), 487--498 
+optimization with combination of Lipschitz bounds. Optim. Methods Softw. \textbf{26}(3), 487--498 
 (2011)
@@
 Evtushenko, Y.G., Posypkin, M.A.: A deterministic approach to global box-constrained 
-optimization. Optim. Lett. 7(4), 819--829 (2013)
+optimization. Optim. Lett. \textbf{7}(4), 819--829 (2013)
 
@@
 
-%\bibitem{Balandin2017}Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 (2017)
+\bibitem{Balandin2017}Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-control and vibroprotection problems. Automation and Remote Control. \textbf{78}(8), 1417--1429 (2017)
 
@@
 Sergeyev, Y.D., Famularo D., Pugliese P. Index Branch-and-Bound Algorithm for Lipschitz 
-univariate global optimization with multiextremal constraints. J. Glob. Optim. 21(3), 317--341 
+univariate global optimization with multiextremal constraints. J. Glob. Optim. \textbf{21}(3), 317--341 
 (2001)
@@
 Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of 
-checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)
+checking for constraints. Comput. Math. Math. Phys. \textbf{42}(9), 1289--1300 (2002)
 
@@
 test functions with known local and global minima for global optimization. ACM Trans. Math. 
-Software 29, 469--480 (2003)
+Software \textbf{29}, 469--480 (2003)
 
@@
 Gergel, V.: An approach for generating test problems of constrained global optimization. 
-LNCS. 10556, 314--319 (2017)
+Lecture Notes in Computer Science. \textbf{10556}, 314--319 (2017)
 
@@
 Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of 
-computationally intensive multiextremal optimization problems. LNCS. 9251, 505--515 (2015)
+computationally intensive multiextremal optimization problems. Lecture Notes in Computer Science. \textbf{9251}, 505--515 (2015)
 
@@
 Barkalov, K., Lebelev, I.: Comparing Two Approaches for Solving Constrained Global 
-Optimization Problems. LNCS. 10556, 301--306 (2017)
+Optimization Problems. Lecture Notes in Computer Science. \textbf{10556}, 301--306 (2017)
 
@@
 Strongin, R.G., Gergel, V.P., Barkalov, K.A.: Parallel methods for global optimization problem 
-solving. Journal of instrument engineering. 52, 25--33 (2009) (In Russian)
+solving. Journal of instrument engineering. \textbf{52}, 25--33 (2009) (In Russian)
 

### 2018-12-11 2016/JOGO2016_rev/JOGO2016_rev.tex

\RequirePackage{fix-cm}
%

%\documentclass{svjour3}                     % onecolumn (standard format)
\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)

%\usepackage[top=3cm, bottom=3cm, right=5cm]{geometry}

%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}

\usepackage{mathptmx}       % selects Times Roman as basic font
%
\usepackage{makeidx}         % allows index generation
\usepackage{multicol}        % used for the two-column index
\usepackage{multirow}

\usepackage{amsmath}

%
%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Journal of Global Optimization}
%
\begin{document}

%\fontsize{12}{13pt}\selectfont

\title{Efficient Multicriterial Optimization \\
Based on Intensive Reuse of Search Information
%On the uniform convergance of Global Search Algorithm when solving in parallel a series of optimization problems
\thanks{This study was supported by the Russian Science Foundation, project No 16-11-10150.}
}

\titlerunning{Efficient Multicriterial Optimization}        % if too long for running head

\author{Victor Gergel         \and
        Evgeny Kozinov 
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{V. Gergel  \at
               Institute of Information Technology, Mathematics and Mechanics\\ Lobachevsky State University of Nizhni Novgorod\\ Nizhni Novgorod, Russia \\
              \email{gergel@unn.ru} 
           \and
           E. Kozinov \at
					Institute of Information Technology, Mathematics and Mechanics\\ Lobachevsky State University of Nizhni Novgorod\\ Nizhni Novgorod, Russia \\
							\email{evgeny.kozinov@itmm.unn.ru} 
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}

This paper proposes an efficient method for solving complex multicriterial optimization problems, for which the optimality criteria may be multiextremal and the calculations of the criteria values may be time-consuming. The approach involves reducing multicriterial problems to global optimization ones through minimax convolution of partial criteria, reducing dimensionality by using Peano curves and implementing efficient information-statistical methods for global optimization. To efficiently find the set of Pareto-optimal solutions, it is proposed to reuse all the search information obtained in the course of optimization. The results of computational experiments indicate that the proposed approach greatly reduces the computational complexity of solving multicriterial optimization problems.

\keywords{ Decision making \and multicriterial optimization \and scalarization \and dimensionality reduction \and global optimization algorithm \and search information \and computational complexity}

\end{abstract}

\section{Introduction} \label{intro}

Multicriterial optimization (MCO) is a field of intensive research and applications -- see, for example, monographs \cite{c5,c6,c7,c8,c26,c30,c49} and reviews of theoretical and practical results in this field \cite{c9,c11,c23,c28,c29,c48}. \par

It should be noted that problems of multicriterial optimization are among the most complex optimization problems. The statement of MCO problems covers many classes of optimization problems, including unconstrained optimization, nonlinear programming, global optimization, etc. In addition, the partial criteria of MCO problems are usually contradictory in nature. As a result, solving the MCO problems consists in finding some compromise (efficient) solutions, which cannot be improved with respect to all partial criteria simultaneously. In the course of optimization, one may need to find a number of different efficient solutions or even the entire set of non-dominated solutions (the Pareto domain). \par

A widely used method for finding the desired solution is to scalarize of the MCO criteria where the partial criteria are combined into a single generalized efficiency criterion -- see, for instance, \cite{c8,c9}. When using such an approach, the convolution coefficients can be considered indicators of the importance of partial criteria -- the greater the coefficient value for some criterion, the greater the contribution of this partial criterion to the single scalar criterion.\par

The need to find several (or the entire set of) efficient solutions increases dramatically the computational complexity in solving MCO problems. Furthermore, it should be noted that the partial criteria can be multiextremal, and the calculations of the criteria values can be time-consuming. In such circumstances, even the finding of just one compromise solution requires a large amount of computations. Thus, finding several (or the whole set of) efficient solutions becomes a problem of huge computational complexity. \par

To overcome the computational complexity of MCO problems, this paper proposes an approach based on the following two main ideas. First, to solve MCO problems, efficient global optimization algorithms will be used, which are based on the information-statistical theory of global optimization \cite{c37,c38}. Second, to accelerate computations, all the search information obtained during the optimization process is totally reused. In general, the reuse of search information will result in a substantial reduction of the amount of computations in searching for the subsequent efficient solutions. \par

Initially, the proposed approach was estimated experimentally in \cite{c47}. In the present work, the developed approach that involves the reuse of the search information has been generalized up to a general scheme for solving multicriterial optimization problems. Within the framework of this scheme, various global optimization algorithms can be used. The efficiency of the proposed scheme has been studied on the example of the information-statistical algorithms, for which the theoretical estimates have been obtained. \par

The rest of the paper is structured as follows. Section 2 states the multicriterial optimization problem and describes the approach for reducing MCO problems to one-dimensional global optimization ones. In Section 3, the proposed approach for solving multicriterial global optimization problems is considered. Section 4 presents the results of numerical experiments. The obtained results are discussed and the directions for further research are considered in the Conclusion.

\section{Problem Statement}  \label{sec:2}
The {\it multicriterial optimization} (MCO) problem can be formulated as follows:

\begin{eqnarray} \label{eq:01}
f(y) = (f_1(y), f_2(y),\dots, f_s(y)) \to \min,  y \in D,
\end{eqnarray}
where
\begin{itemize}
\item	$y = (y_1, y_2, \dots , y_N)$ is a vector of varied parameters,
\item	$N$ is the MCO problem dimensionality,
\item	$D$ is the search domain, which is a $N$-dimensional box
\begin{eqnarray*}
D  = \{ y \in R^N: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}
\end{eqnarray*}
\end{itemize}
with the given boundary vectors $a$ and $b$. Without a loss of generality, it is supposed that the values of the partial criteria are not negative and their decrease corresponds to an increase in the solution efficiency. \par

Any efficient point can be considered as {\it partial solution} of the MCO problem. In general, solving a MCO problem may require finding the whole set of the Pareto-optimal points $PD \subseteq D$ ({\it the complete solution of the MCO problem}). \par

In this paper, the MCO problems are considered in relation to the most complicated decision making problems, in which the partial criteria $f_i (y)$, $1 \leq i \leq s$, can be multiextremal, and obtaining the criteria values at the search domain points $y \in D$  may require time-consuming computations. We also suppose that the partial criteria $f_i (y)$, $1 \leq i \leq s$, satisfy the Lipschitz condition

\begin{eqnarray} \label{eq:02}
|f_i(y') - f_i(y'')| \leq L_i \|y' - y''\|, y', y'' \in D, 1 \leq i \leq s. 
\end{eqnarray}

Satisfying the Lipschitz condition allows us to compute numerical estimates of the possible partial criteria values based on their calculated values on a finite set of points within the search domain $D$. \par

The proposed approach to solving MCO problems is based on the following two key techniques. \par

	{\bf 1. Scalarization.} One of the widely used approaches in addressing MCO problems is the scalarization technique, which employs various convolution schemes of the partial criteria $f_i (y)$, $1 \leq i \leq s$, to obtain a scalar generalized criterion $F(\lambda,y)$ -- see, for instance, \cite{c8,c9}. This approach makes it possible to reduce the solution of problem (\ref{eq:01}) to the solution of a family of nonlinear optimization subproblems

\begin{eqnarray} \label{eq:03}
\min F(\lambda,y), y \in D 
\end{eqnarray}
where $\lambda = (\lambda_1  ,\lambda_2,\dots,\lambda_s)$ is the vector of coefficients used for formulating the scalar criterion $F(\lambda,y)$. It is assumed that in addition to (\ref{eq:02}) the scalar criterion $F(\lambda,y)$ also satisfies the Lipschitz condition

\begin{eqnarray} \label{eq:04}
|F(\lambda, y') - F(\lambda, y'')| \leq L \|y' - y''\|, y', y'' \in D, 1 \leq i \leq s. 
\end{eqnarray}

In this paper, it is proposed to use the minimax convolution scheme in the following form:

\begin{eqnarray} \label{eq:05}
F(\lambda,y) = \max_{1 \leq i \leq s}(\lambda_i f_i(y)), 
\end{eqnarray}
with the given coefficients $\lambda_i$, $1 \leq i \leq s$, that should belong to the set

\begin{eqnarray*}
   \Lambda \subset R^s:\sum_{i=1}^s{\lambda_i} = 1, \lambda_i \geq 0, 1 \leq i \leq s.
\end{eqnarray*}

These coefficients $\lambda_i$, $1 \leq i \leq s$, can be considered indicators of partial criteria importance -- the greater the value of some criterion coefficient $\lambda_i$, the greater the contribution of this partial criterion $f_i$ in the scalar criterion $F(\lambda,y)$. \par

An important property of the minimax convolution scheme is the necessity and sufficiency of this approach for solving MCO problems: the result of minimizing $F(\lambda,y)$ from (\ref{eq:05}) yields an efficient solution\footnote{More precisely, minimizing $F(\lambda,y)$ can result in obtaining a weakly-efficient solution (a set of weakly-efficient points includes the Pareto domain). This can be corrected by adding an additional correction value in (\ref{eq:05}) -- see, for example, \cite{c19,c29}.}  of a MCO problem. And, vice versa, every efficient solution of a MCO problem can be obtained through minimizing $F(\lambda,y)$ for the corresponding convolution coefficients $\lambda_i$, $1 \leq i \leq s$, -- see, for instance, \cite{c19,c29}.\par

It should be pointed out that in the process of solving a MCO problem, the concept of the importance of partial criteria may change, which may require solving several subproblems (\ref{eq:05}) using different coefficients $\lambda \in \Lambda$. As a result, the possibility of finding a number (or the whole set) of efficient solutions with adequate computational costs becomes the key issue in solving MCO problems.\par

{\bf 2. Dimensionality reduction.} To decrease computational complexity of solving multidimensional optimization problems, the dimensionality reduction can be applied by using {\it Peano curves or mappings} $y(x)$, that map unambiguously the interval $[0,1]$ on a $N$-dimensional hypercube $D$ \cite{c35,c37,c38}. \par

Following such a reduction scheme, the initial multidimensional global optimization subproblem (\ref{eq:03})-(\ref{eq:05}) is reduced to the one-dimensional subproblem:

\begin{eqnarray} \label{eq:06}
F(\lambda, y(x^*)) = \min_{x \in [0,1]}{F(\lambda, y(x))}.
\end{eqnarray}

It should be mentioned that the reduced function $F(\lambda, y(x))$ satisfies the uniform H\"{o}lder condition, i.e.

\begin{eqnarray} \label{eq:07}
|F(\lambda,y(x')) - F(\lambda,y(x''))| \leq H |x' - x''|^{1/N}, x', x'' \in [0,1], 
\end{eqnarray}
where the constant $H$ is determined by the relation $H = 2L \sqrt{N+3}$, where $L$ is the Lipschitz constant from (\ref{eq:04}), and $N$ is the dimensionality of the optimization problem (\ref{eq:01}). \par

The computational scheme for computing the values of the function $F(\lambda, y(x))$ obtained through the dimensionality reduction is as follows:
\begin{itemize}
	\item The optimization algorithm solves the reduced one-dimensional function $F(\lambda, y(x))$,
	\item Once the next trial point $x \in [0,1]$ is selected, a multidimensional image $y \in D$ using the mapping $y(x)$, is calculated,
	\item At the point $y \in D$ the value of the function $F(\lambda, y)$ is calculated,
	\item The calculated value $z = F(\lambda, y)$ will be used further as the value of the reduced one-dimensional function $F(\lambda, y(x))$ at the point $x \in [0,1]$.
\end{itemize}

The algorithms for constructing the numerical approximations of Peano curves are given in \cite{c38}.

\section{Methods for Solving Multicriterial Global Optimization Problems}  \label{sec:3}
As it was mentioned above the scalarization approach makes it possible to use efficient global optimization methods for solving MCO problems. Finding numerical estimates for the globally optimal points is usually provided by a coverage of the search domain D -- see, for instance, \cite{c12,c22,c24,c25,c27,c31,c35,c37,c38,c39,c41,c44,c51,c52}. As a result, the computational complexity of solving the global optimization problems is very high even for a fairly small number of varied parameters (the dimensionality of the problem). A dramatic reduction of the required computations can be achieved if the computing grids generated for covering the search domain are non-uniform, with the denser optimization points only in the vicinity of the desired globally optimal solutions. Such non-uniform coverage can be constructed by performing adaptive computations. However, this is possible if and only if all the search information (the previous optimization points and the values of the optimized function at these points) obtained through the calculations is taken into account for selecting each iteration point.

\subsection{Reusing Search Information within Multicriterial Global Optimization}
As follows from the above discussion, to improve the efficiency of the global optimization algorithms, the search information (the trial points and the corresponding values of the objective function) should be accumulated and used intensively within optimization procedures. Search information in conjunction with the Lipschitz condition allows us to calculate the estimates of the possible values of the optimized function. Besides, search information makes possible the adaptive selection of the points for the subsequent global search iterations. And contrariwise, a reduction in the amount of the search information usually leads to a significant increase in the number of iterations required to meet the stopping condition. \par

The search information obtained during the calculations may be represented as the set

\begin{eqnarray} \label{eq:08}
A_k = \{(x_i, z_i, v_i)^T: 1 \leq i \leq k \}
\end{eqnarray}
where
\begin{itemize}
	\item $x_i$, $1 \leq i \leq k$, are the points of the executed global optimization iterations; the points should be located in the increasing order;
	\item $z_i$, $1 \leq i \leq k$, are the scalar criterion values for the optimization subproblem (\ref{eq:06}) at the points $x_i$, $1 \leq i \leq k$, i.e.:
\begin{eqnarray*}
z_i = F(\lambda,x_i), 1 \leq i \leq k;
\end{eqnarray*}

	\item $v_i$, $1 \leq i \leq k$, are the values of the vector criterion $f(y(x))$ of the initial multicriterial optimization problem (\ref{eq:01}) calculated at the points  $x_i$, $1 \leq i \leq k$, i.e.:
\begin{eqnarray*}
v_i=(v_i^1,v_i^2,\dots,v_i^s ), v_i^j=f_j (y(x_i)), 1 \leq j \leq s, 1 \leq i \leq k.
\end{eqnarray*}

\end{itemize}

It should be pointed out that the search information $A_k$ contains the calculated partial criteria values of the initial MCO problem. The availability of such information allows the current values $z_i$, $1 \leq i \leq k$, to be recalculated to those for the next optimization problem (\ref{eq:06}) with any new value of the convolution coefficients $\lambda_j'$, $1 \leq j \leq s$, i.e.

\begin{eqnarray} \label{eq:09}
z_i' = \max{(\lambda_j' v_i^j, 1 \leq j \leq s)},1 \leq i \leq k,
\end{eqnarray}
without time-consuming calculations of the partial criteria values $f_i (y(x))$, $1 \leq j \leq s$. The number of computational operations required for recalculating the search information can be evaluated as $M=2sk$, that can be considered negligible compared with the time-consuming computations of the partial criteria values.\par

As a result of recalculating the values $z_i$, $1 \leq i \leq k$, the search information $A_k$ from (\ref{eq:08}) will be readily available to use for solving the subsequent optimization subproblems (\ref{eq:06}). Thus, the global search can be continued instead of being restarted. Available search information will help reducing the number of global search iterations required to satisfy the stopping condition. It is expected that the amount of search information could reach such a large scale that solving each subsequent subproblem would require a minimum number of additional calculations (see the results of the numerical experiments in Section 4). \par

In particular, the availability of the search information allows computing the numerical estimates of the H\"{o}lder constant $H$ from (\ref{eq:07})
\begin{equation} \label{eq:10}
m = 
 \begin{cases}
   r M, & M > 0 \\
   1,   & M = 0
 \end{cases} ,
 M = \max_{1 < i \leq k} {\frac{|z_i - z_{i-1}|} {\varrho_i} }
\end{equation}
as the relative differences of the values of the objective function $z_i=F(\lambda,y(x_i))$, $1 \leq i \leq k$ on the set of points $x_i$,$1\leq i \leq k$. Hereafter $\varrho_i=\sqrt[N]{x_i - x_{i-1}}$, $1 < i \leq k$. The constant $r$, $r>1$ is the {\it reliability parameter} of the estimate of the constant $H$.

\subsection{General Computational Scheme with Reusing Search Information for Solving MCO Problems}
As a result of the above discussion the general computational scheme which provides reusing the search information for solving MCO problems can be proposed as follows.\par

Prior to beginning the computations, it is necessary to form the set $\Delta_q$ consisting of the values of the vectors $\lambda \in \Lambda$ from (\ref{eq:05})
\begin{eqnarray} \label{eq:11}
\Delta_q = \{\lambda_i \in \Lambda: 1 \leq i \leq q\},
\end{eqnarray}
for each of which the problems (\ref{eq:06}) should be solved in order to find the efficient solutions of the MCO problem. The set $\Delta_q$ can be modified in the course of computations if necessary.\par

The computational scheme for solving the set of problems (\ref{eq:06}) for the corresponding vectors $\lambda \in \Delta_q$ is an iteration procedure, where at every iteration the following operations are performed.
\begin{enumerate}
	\item Select the next vector $\lambda \in \Delta_q$ from the set $\Delta_q$ thus defining a new problem (\ref{eq:06}); the chosen vector $\lambda$ should be excluded from the set $\Delta_q$.
	\item Solve the formulated problem (\ref{eq:06}) using any global optimization algorithm; the algorithm should utilize the search information $A_k$ from (\ref{eq:08}) in the planning of the executed global search iterations; the additional search information obtained in the course of computations should be included into the set $A_k$.
	\item Analyze the obtained results of solving the current problem (\ref{eq:06}); modify (reduce or extend) the content of the set $\Delta_q$. if necessary.
	\item Terminate the computations if the set $\Delta_q$ is empty, otherwise go to Step 1, and the computations should be continued further.
\end{enumerate}

Each global optimization problem (\ref{eq:06}) is solved by the optimization method with the required accuracy in accordance to the MAGS convergence properties independently on the reused search information. This statement is confirmed also by the numerical results (see Section 4, Tables \ref{tab:03} - \ref{tab:05}), where the number of problems solved with the required accuracy are almost the same, and the indicators of completeness and uniformity of the approximation of the Pareto domain at approximately equal allowed number of  iterations are higher when reusing the search information (see Table \ref{tab:05}). The benefit of the reuse of the search information is a possible decrease in the number of executed iterations required to provide the required accuracy.\par

\subsection{Multidimensional Algorithm for Multicriterial Global Optimization}
According to the proposed approach, MCO problems can be solved by using numerous efficient one-dimensional optimization methods. However, it should be noted that one-dimensional optimization algorithms require certain generalizations when applied to multidimensional problems. Moreover, the applied algorithms should use the search information in accordance with the computational scheme proposed in Section 3.2.\par

The main results of using Peano curves for the dimensionality reduction are obtained within the framework of the information-statistical optimization theory. This theory has provided the basis for developing a large number of efficient multiextremal optimization methods -- see, for instance, \cite{c1,c2,c3,c15,c16,c17,c18,c21,c32,c33}, etc. \par

The proposed approach employs the {\it Multidimensional Algorithm of Global Search} (MAGS) \cite{c37,c38} for solving one-dimensional multiextremal optimization subproblems (\ref{eq:06}). This method has a strong theoretical basis and has been proved experimentally to be highly efficient as compared to other global search algorithms (see also the results of computational experiments in Section 4).

The general computational scheme of MAGS is given below in summary form.\par

Let $k$, $k>1$ be the number of executed optimization iterations, the values of the objective function $F(\lambda,y(x))$ (hereinafter, this procedure is called the {\it trial}) have been computed for all previous iterations, and the obtained search information is represented in the form of $A_k$ from (\ref{eq:08}). The trial points of the next $(k+1)^{th}$ iteration are determined by the following rules.

{\it Rule 1.} Compute the {\it characteristic} $R(i)$ for each interval $(x_{i-1}, x_i)$, $1 < i \leq k$ from $A_k$

\begin{eqnarray} \label{eq:12}
R(i) = \varrho_i + \frac{(z_i - z_{i-1})^2}{m^2\varrho_i} - 2 \frac{(z_i + z_{i-1})}{m} , 1 < i \leq k, 
\end{eqnarray}
where $m$, $z_i$, $1 \leq i \leq k$ and $\varrho_i$, $1 < i\leq k$ are defined in (\ref{eq:10}).\par

{\it Rule 2.} Determine the interval $(x_{t-1}, x_t)$ with the highest characteristic $R(t)$, i. e.
\begin{eqnarray} \label{eq:13}
R(t) = \max_{1 < i \leq k} {R(i)}. 
\end{eqnarray}

{\it Rule 3.} Compute the trial point $x^{k+1}$ of the next iteration within the interval $t$ from (\ref{eq:13}):
\begin{eqnarray*}
x^{k+1} = \frac{x_t + x_{t-1}}{2} - sign(z_t - z_{t-1})\frac{1}{2r} \left[\frac{r|z_t - z_{t-1}|}{m} \right]^N
\end{eqnarray*}

The {\it stopping condition} is defined by the inequality
\begin{eqnarray} \label{eq:14}
\varrho_t \leq \epsilon 
\end{eqnarray}
which should be checked for the interval $t$ from (\ref{eq:13}). The value $\epsilon > 0$ is the predefined {\it accuracy} of solving the global optimization problem. If the stopping condition is not fulfilled, the iteration index $k$ is incremented by 1, and the execution of the algorithm should be continued. \par

The lowest computed value of the objective function $F(\lambda,y(x))$ is taken as the current estimate of the global minimum value, i. e.:

\begin{eqnarray} \label{eq:15}
z_k^* = \min_{1 \leq i \leq k} {z_i}.
\end{eqnarray}

It can be noted that the characteristics $R(i)$, $1 < t \leq k$ from (\ref{eq:12}) can be considered as some measures of importance of the intervals with respect to the location of the global minimum point in these intervals.\par

The MAGS algorithm applied for solving the MCO problems and combined with the computational scheme given in 3.2 will be further referred to as the {\it Multidimensional Algorithm of Multicriterial Global Search} (MAMGS).

\subsection{Theoretical Analysis of the Multicriterial Global Optimization Algorithm}
For the sake of simplicity, a theoretical analysis is implemented for the Algorithm of Global Search (AGS) which is the one-dimensional version of MAGS. The proven properties of AGS are also true for MAGS and MAMGS. In this case, a reduced subproblem of multiextremal optimization (6) can be stated as follows:
\begin{eqnarray} \label{eq:16}
\phi(x^*) = \min {\phi(x)}, x \in[0,1] 
\end{eqnarray}
where $\phi(x) = F(\lambda,x)$ from (\ref{eq:06}).\par

First and foremost, the AGS algorithm is stable. The following theorem holds true. \par

{\bf Theorem 1 \cite{c37,c38}.} Suppose that along with minimizing the multiextremal function $\phi(x)$ the AGS is also used for optimizing the function $\psi(x)$, which  differs from $\phi(x)$ by some limited variation, i.e.
\begin{eqnarray} \label{eq:17}
|\psi(x) - \phi(x)| \leq \delta, x \in[0,1] 
\end{eqnarray}
and suppose that in the process of minimization $\psi(x)$ the stopping condition (\ref{eq:14}) is satisfied with the accuracy $\epsilon > 0$. Then
\begin{eqnarray} \label{eq:18}
|z_k^*-\phi(x^*)| \leq \frac{\alpha L \epsilon}{2} + \delta, 
\end{eqnarray}
if  for the value $m$ from (\ref{eq:10}) and some $ \alpha> 1$ the inequality
\begin{eqnarray} \label{eq:19}
m \geq L\left( 1+\sqrt{ \frac{\alpha + 4 \beta(1 + \beta)}{\alpha - 1}} \right),
\end{eqnarray}
is satisfied where
\begin{eqnarray} \label{eq:20}
\beta = \frac{2 r \delta}{L\epsilon(r-1)}
\end{eqnarray}
and $z_k^*$ from (\ref{eq:15}), $\phi(x^*)$ from (\ref{eq:16}), $L$ from (\ref{eq:04}). \par

The statement of Theorem 1 can be used to analyze the results of solving the series of optimization subproblems (\ref{eq:16}). Suppose that in the course of solving the MCO problem, we proceeded from solving the subproblem (\ref{eq:16}) with the convolution coefficients $\lambda' \in \Lambda$ from (\ref{eq:05}) to solving the subproblem (\ref{eq:16}) with the coefficients $\lambda'' \in \Lambda$. Let us estimate the maximum variation of the coefficient values  $\lambda_i$, $1 \leq i \leq s$ obtained as a result of such a transition
\begin{eqnarray} \label{eq:21}
\Delta \lambda_{max}=\max_{1 \leq i \leq s}{|\lambda_i'-\lambda_i''|}.
\end{eqnarray}

Under these conditions, the maximum variation of the objective function values after stating the new subproblem (\ref{eq:16}) is bounded by the value
\begin{eqnarray} \label{eq:22}
\Delta \phi_{max}=\max{\{\Delta \lambda_{max} f_i^{max}, 1 \leq i \leq s \},}
\end{eqnarray}
where
\begin{eqnarray*} 
f_i^{max} = \max { \{ f_i (y), y \in D \}, 1 \leq i \leq s  }.
\end{eqnarray*}

Using the obtained estimates, Theorem 1 can be reformulated in relation to the MCO problems.\par

{\bf Corollary 1.} When switching from solving the optimization subproblem (\ref{eq:16}) with the convolution coefficients $\lambda' \in \Lambda$ from (\ref{eq:05}) to solving the next subproblem (\ref{eq:16}) with the coefficients $\lambda'' \in \Lambda$, the value of the function $F(\lambda'',x)$ at the point of the global minimum differs from the estimated minimum value $z_k^*$ from (\ref{eq:15}) obtained by minimizing the function $F(\lambda',x)$, by no more than the bounded value 
\begin{eqnarray} \label{eq:23}
|z_k^*-F(\lambda'',x^*)| \leq \frac{\alpha L \epsilon}{2} + \delta, \alpha>1, 
\end{eqnarray}
where $\delta=\Delta \phi_{max}$ from (\ref{eq:21})-(\ref{eq:22}) and the conditions (\ref{eq:19})-(\ref{eq:20}) are satisfied.\par

This statement means that if the deviation $\Delta \phi_{max}$ from (\ref{eq:22}) of the estimates of the minimum values of $F(\lambda'',x)$ is acceptable, then solving the subproblem (\ref{eq:16}) with the new convolution coefficients $\lambda'' \in \Lambda$ does not require any additional global search iterations -- the estimates of the minimum value of the function $F(\lambda'',x)$ can be obtained in accordance with (\ref{eq:23}) using the values $z_i$, $1 \leq i \leq k$, located within the search information $A_k$ from (\ref{eq:08}).\par

Further analysis is related to the distribution of iteration points within the search domain. The density of trials $P_{ab}$ in the subinterval $[a,b] \subset [0,1]$ can be defined as the ratio of the number $N_{ab}$ of executed trial points $x_i$, $1 \leq i \leq k$, belonging to the interval $[a,b]$, to its length, i.e.

\begin{eqnarray} \label{eq:24}
P_{ab} = \frac{N_{ab}}{b-a}.
\end{eqnarray}
The following theorem holds true.\par

{\bf Theorem 2 \cite{c37,c38}.} Suppose that the objective function $\phi(x)$, $x \in [0,1]$, from (\ref{eq:16}) is the Lipschitzian function with the constant $L$ and 
\begin{eqnarray} \label{eq:25}
\phi (x) \geq \phi(x^*)+\Delta, x \in [a,b] \subset [0,1],
\end{eqnarray}
where $x^*$ is the global minimum point of the function $\phi(x)$ and
\begin{eqnarray*}
b - a > 2 \Delta / L.
\end{eqnarray*}
Then
\begin{eqnarray*}
P_{ab} < 3 m /2 \Delta,
\end{eqnarray*}
if the value $m$ satisfies the inequality $m \geq 2L$, $m$ from (\ref{eq:10}), $P_{ab}$ from (\ref{eq:24}).\par

Theorem 2 can be used to estimate the density of the executed trials for solving the subproblems of the family (\ref{eq:16}).

{\bf Corollary 2.} Suppose that in solving the next subproblem (\ref{eq:16}), the conditions of Theorem 2 are satisfied, all  the previous trial points $x_i$, $1 \leq i \leq k$, from the search information $A_k$ are taken into account and the density $p_{ab}$ of a given interval $[a,b]$ has been obtained. Then the number of additional global search iterations  $n_{ab}'$ within the interval [a,b] executed for solving the next subproblem (\ref{eq:16}) cannot exceed the value 
\begin{eqnarray*}
n_{ab}' \leq (3 m / 2 \Delta - p_{ab} )(b-a)  .
\end{eqnarray*}

This statement means that, in solving the next subproblems of the family (\ref{eq:16}), when a sufficiently high trial density is obtained, the number of additional iterations within the intervals located far enough from the global minimum point will be constantly decreasing (up to the total absence of these additional iterations).\par

As a result of the theoretical analysis, two different schemes for forming various convolution coefficients $\lambda \in \Lambda$ can be proposed, at the use of which the conditions of Consequence 1 of Theorem 1 will be satisfied. In the first scheme, initially a sparse grid can be built on the set $\Lambda$, so that when setting new values of $\lambda$ later, the close values could be always found in the grid built in Λ earlier. In the second scheme, new values of $\lambda$ could be chosen as a result of small variations of the values used earlier (this method is used in the practical solving the MCO problems often, when the decision maker wants to refine the estimates of the efficient solutions obtained earlier). It is worth noting that the reuse of the search information could be useful even at essential differences in the chosen values of the coefficients $\lambda$ (see Section 4).

\section{Results of computational experiments}
As it was mentioned above, the proposed approach was developed on the basis of the minimax scalarization technique (\ref{eq:05}) and by reducing MCO problems to a series of global optimization problems (\ref{eq:06}). As a result, the numerical evaluation of the approach efficiency should focus on a comparison of the MAGS algorithm with other multiextremal optimization methods. Such a comparisons have been made in a sufficiently large number of works -- see, for instance, \cite{c1,c2,c3,c16,c18,c20,c25,c31,c32,c33,c35}. These studies show, in particular, that MAGS is one of the most efficient multiextremal optimization methods.\par

The computational experiments have been performed using the Lobachevsky supercomputer at the State University of Nizhny Novgorod (operating system -- CentOS 6.4, supercomputer management system -- SLURM). Each computational node of the supercomputer had 2 Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM processors. Each CPU had 8-cores (that is, 16 CPU cores were available at the computing node). To generate the executable program code, the Intel C++ Compiler 14.0.2 was used.\par

The experiments were executed as follows. At the beginning, a comparison of the proposed approach with a number of other multicriterial optimization algorithms was performed. Then the MAMGS efficiency is evaluated by solving a large set of bi-criteria two-dimensional optimization problems. Finally, the experiments were executed for more complex optimization problems with 5 varied parameters and 5 partial criteria.\par

In all numerical experiments considered below, in order to construct the Pareto domain approximation with the use of MAMGS, the set of subproblems (\ref{eq:06}) with different convolution coefficients $\lambda$ from (\ref{eq:05}) distributed uniformly in $\Lambda$ was solved. \par

Besides, every time after solving this set of subproblems, the additional non-dominated points have been selected from the obtained search information -- this approach allows expanding the number of the efficient points making up the Pareto domain approximation. \par

In the first series of experiments the proposed approach is compared with the following five multicriterial optimization methods:
\begin{itemize}
	\item The Monte-Carlo (MC) method, where the trial points are selected within the search domain $D$ randomly and uniformly,
	\item The genetic algorithm SEMO from the PISA library \cite{c4,c10},
	\item The non-uniform coverage (NUC) method \cite{c10},
	\item The bi-objective Lipschitz optimization (BLO) method \cite{c45},
	\item The MAMGS method proposed herein.
\end{itemize}

To solve MCO problems, the compared methods should compute some numerical approximation $PDA \subseteq PD$ of the Pareto domain $PD \subseteq D$.\par

For the first three algorithms, the numerical results were used from \cite{c10}. The results of the BLO method were presented in \cite{c45}. \par

To evaluate the quality of the approximation, the completeness and uniformity of the Pareto domain coverage were compared by using the following two indicators \cite{c10,c46}:
\begin{itemize}
	\item The {\it hypervolume index} (HV) defined as the total size of subdomains of the objective criterion values, dominated by the approximation points of the Pareto domain. This indicator's value can be obtained as the total volume of crossing boxes, the vertices of which are the approximation points of the Pareto domain and the reference point $\hat{f}$(see Figure \ref{fig:01}).  As a reference point in the experiments conducted, we used the vector value:
\begin{eqnarray*}
 \hat{f}=(\hat{f}_1,\dots,\hat{f}_s):\hat{f}_i=\max_{1 \leq i \leq s} {\hat{f}_i(y), y \in D}.
\end{eqnarray*}

This indicator characterizes the Pareto domain approximation completeness (the higher the value, the more complete the Pareto domain coverage).

	\item The {\it distribution uniformity index} (DU) of the points from the Pareto domain approximation, defined by the expressions:
\begin{equation*}
\begin{matrix}
DU=\sum_{i=1}^p{(d_i-d)^2} , d=\dfrac{ \sum_{i=1}^p{d_i} }{p}, \\
d_i=\min {\varrho(y_i,y)}, y \in PDA, y_i \neq y, 1 \leq i \leq p,
\end{matrix}
\end{equation*}
where $p$ represents the number of points in the Pareto domain approximation $PDA$. This indicator characterizes the Pareto domain approximation uniformity (the lower the value, the more uniform the Pareto domain coverage).
\end{itemize}

\begin{figure}
\center{\includegraphics[scale=.5]{fig_01}}
\caption{An example of the hypervolume index (HV) estimate}
\label{fig:01}
\end{figure}

The first benchmark problem is stated as follows \cite{c10}:
\begin{equation}\label{eq:26}
f_1 (y)=(y_1-1) y_2^2+1, f_2 (y)=y_2, 0 \leq y_1,y_2 \leq 1.
\end{equation}

For MAMGS, the following parameters were used: the reliability parameter $r=2$, the accuracy $\epsilon = 0.06$, and 100 subproblems (\ref{eq:06}) have been solved. The reference point was (1,1).\par

The results of the experiments are presented in Table \ref{tab:01}. The column ''Iterations'' shows the total number of the optimization iterations executed by the method (this is the same as the total number of the iteration points at which the objective function values are computed). The column ''PDA points'' demonstrates the number of points included in the numerical approximation of the Pareto domain. The efficiency indicators of the calculated Pareto domain approximations are given in the columns ''HV'' and ''DU''.

\begin{table}[t]
\centering
\caption{Numerical results of the methods compared for the benchmark problem  (\ref{eq:26})}
\label{tab:01}
\begin{tabular}{cllll}
\hline
\textbf{Method} & \textbf{Iterations} & \textbf{PDA points} & \textbf{HV} & \textbf{DU} \\ \hline
MC              & 500                 & 67                  & 0.300       & 1.277       \\
SEMO            & 500                 & 104                 & 0.312       & 1.116       \\
NUC             & 515                 & 29                  & 0.306       & 0.210       \\
BLO             & 498                 & 68                  & 0.308       & 0.175       \\
MAMGS           & 390                 & 90                  & 0.317       & 0.094       \\ \hline
\end{tabular}
\end{table}

To illustrate the obtained results, Figure \ref{fig:01} shows the Pareto domain approximation obtained by using the MAMGS method.

\begin{figure}
\center{\includegraphics[scale=.4]{fig_02}}
\caption{The Pareto domain approximation (a) and iteration points (b) executed by MAMGS}
\label{fig:02}
\end{figure}

Further, the second benchmark problem has been solved \cite{c10}:
\begin{equation}\label{eq:27}
f_1 (y)=y_1, f_2 (y)=\min {( |y_1-1|,1.5-y_1)}+y_2+1, 0 \leq y_1,y_2 \leq 2.
\end{equation}

For MAMGS, the following parameters were used: the reliability parameter $r=2$, the accuracy $\epsilon = 0.06$, and 100 subproblems (\ref{eq:06}) have been solved. The reference point was (2,3).
The results of the experiments are presented in Table 2 (the values of the DU indicator are not available in \cite{c10,c45}).

\begin{table}[t]
\centering
\caption{Numerical results of the methods compared for the benchmark problem (\ref{eq:27})}
\label{tab:02}
\begin{tabular}{clll}
\hline
\textbf{Method} & \textbf{Iterations} & \textbf{PDA points} & \textbf{HV} \\ \hline
MC              & 500                 & 22                  & 3.38        \\
SEMO            & 500                 & 221                 & 3.27        \\
NUC             & 490                 & 36                  & 3.42        \\
BLO             & 435                 & 92                  & 3.61        \\
MAMGS           & 380                 & 92                  & 3.59        \\ \hline
\end{tabular}
\end{table}

As shown by the experimental results, MAMGS has an evident advantage over a number of other multicriterial optimization methods, even in solving relatively simple MCO problems.\par

It should be noted that the conditions for this experiment did not correspond to the initial assumptions about the properties of the multicriterial optimization problems: the partial criteria are neither multiextremal nor time-consuming.\par

In the second series of experiments, MAMGS was compared with four other MCO methods considered in \cite{c50}, namely: 
\begin{itemize}
	\item The linear combination method (LCM),
	\item The Multi-Objective Genetic Algorithm (MOGA),
	\item The global criterion method (GCM),
	\item The $\epsilon$-constraint method (ECM).
\end{itemize}

The benchmark problems and the numerical results of these methods are taken form \cite{c50}.\par

The first benchmark problem is stated as follows:
\begin{equation}\label{eq:28}
 \begin{cases}
   f_1 (y)=1-\exp \left( -\sum_{i=1}^3{(y_i - 1/ \sqrt{3})^2} \right), \\
   f_2 (y)=1-\exp \left( -\sum_{i=1}^3{(y_i + 1/ \sqrt{3})^2} \right), \\
 \end{cases}
-4 \leq y_1,y_2 \leq 4.
\end{equation}

For MAMGS, the following parameters were used: the reliability parameter $r=1.1$, the accuracy $\epsilon = 0.04$, and 50 subproblems (\ref{eq:06}) have been solved. \par

Because the numerical results in \cite{c50} are given in the graphic form, the results of the experiment are given in the same way -- see Figure \ref{fig:03}. In addition, the number of executed iterations can be indicated: $LCM=258$, $MOGA=2500$, $GCM=286$, $ECM=257$. For MAMGS the number of executed iterations is 283.

\begin{figure}
\center{\includegraphics[scale=.5]{fig_03}}
\caption{The Pareto domain approximations calculated by the compared methods for the benchmark problem (\ref{eq:28})}
\label{fig:03}
\end{figure}

The second benchmark problem is stated as follows:
\begin{equation}\label{eq:29}
f_1 (y)=y_1,f_2 (y)=(1+10y_2 ) \left[1-(y_1/(1+ 10 y_2 ))^2-y_1/(1+10y_2 ) sin⁡(8 \pi y_1) \right],
\end{equation}
where $0 \leq y_1,y_2 \leq 1$.\par

For MAMGS, the following parameters were used: the reliability parameter $r=2$, the accuracy $\epsilon = 0.01$, and 50 subproblems (\ref{eq:06}) have been solved.\par

The numerical results are given in the graphic form -- see Figure \ref{fig:04}. In addition the number of executed iterations can be indicated: $LCM=875$, $MOGA=3000$, $GCM=427$, $ECM=575$. For MAMGS the number of executed iterations is 277.

\begin{figure}
\center{\includegraphics[scale=.5]{fig_04}}
\caption{The Pareto domain approximations calculated by the compared methods for the benchmark problem (\ref{eq:29})}
\label{fig:04}
\end{figure}

In the third series of experiments, the bi-criterial two-dimensional optimization problems have been solved, i.e. $N=2$, $s=2$. To obtain the MCO benchmark problems, the multiextremal functions were generated using the GKLS-generator \cite{c14}. This generator generates the multiextremal optimization problems with a priori known properties: the number of local minima, the size of their attraction domains, the point of global minimum, and the value of the functions therein, etc.\par

For MAMGS the following parameters were used: the reliability parameter $r=4.5$, the accuracy $\epsilon = 0.01$. A total of 100 MCO problems were solved, for each of them 50 global optimization subproblems (\ref{eq:06}) were solved with different convolution coefficients $\lambda$, distributed in $\Lambda$ from (\ref{eq:06}) uniformly. \par

The experimental results are shown in Table \ref{tab:03}. The columns ''10'',\dots,''50'' show the average number of optimization iterations executed by the method for solving the indicated set of the subproblems (\ref{eq:06}).

\begin{table}[t]
\centering
\caption{Experimental results for solving two-dimensional bi-criterial problems}
\label{tab:03}
\begin{tabular}{ccccccc}
\hline
Computational results                                                                                  &           & \multicolumn{5}{c}{Number of subprolems being solved}                        \\ \cline{3-7} 
(average per subproblem)                                                                               &           & 10           & 20            & 30            & 40            & 50            \\ \hline
\begin{tabular}[c]{@{}c@{}}Computations without \\ using search information\end{tabular}               &           & 647.7        & 817.8         & 840.6         & 882.0         & 835.2         \\
\begin{tabular}[c]{@{}c@{}}Number of subproblems solved \\ with the required accuracy\end{tabular}     &           & 99.4\%       & 99.1\%        & 98.6\%        & 98.7\%        & 98.8\%        \\
\hline
\begin{tabular}[c]{@{}c@{}}Computations using \\ search information\end{tabular}                       &           & 94.1         & 87.3          & 89.9          & 82.7          & 67.1          \\
\begin{tabular}[c]{@{}c@{}}Number of subproblems solved \\ with the required accuracy\end{tabular}     &           & 99.4\%       & 98.9\%        & 98.6\%        & 98.7\%        & 98.9\%        \\
\hline
\textbf{\begin{tabular}[c]{@{}c@{}}Reduction in the number of \\ optimization iterations\end{tabular}} & \textbf{} & \textbf{6.9} & \textbf{9.4}  & \textbf{9.4}  & \textbf{10.7} & \textbf{12.5} \\ \hline
\end{tabular}
\end{table}

These results show that if the method uses the search information then it can reduce the executed optimization iterations in solving multicriterial optimization problems up to 12.5 times. Moreover, as it has been noted previously, solving each subsequent scalar subproblem (\ref{eq:06}) to find another partial solution of the MCO problem requires a permanently decreasing number of optimization iterations. For  more explicit demonstration of this key property of the proposed approach, Table \ref{tab:04} shows the results of experiments separated by the sequence of solved subproblems (\ref{eq:06}); each group includes 10 scalar subproblems (\ref{eq:05}).

\begin{table}[t]
\centering
\caption{Average number of executed iterations per  a subproblem within successive groups including 10 subproblems}
\label{tab:04}
\begin{tabular}{cccccc}
\hline
\begin{tabular}[c]{@{}c@{}}Number of problems \\ included in a group\end{tabular}                      & 1-10         & 11-20         & 21-30        & 31-40         & 41-50          \\ \hline
\begin{tabular}[c]{@{}c@{}}Computations without \\ using search information\end{tabular}               & 647.7        & 987.8         & 886.2        & 1,006.1       & 648.2          \\
\begin{tabular}[c]{@{}c@{}}Computations using \\ search information\end{tabular}                       & 94.1         & 80.6          & 94.9         & 61.1          & 4.6            \\
\textbf{\begin{tabular}[c]{@{}c@{}}Reduction in the number\\  of optimization iterations\end{tabular}} & \textbf{6.9} & \textbf{12.3} & \textbf{9.3} & \textbf{16.5} & \textbf{139.4} \\ \hline
\end{tabular}
\end{table}

The experimental results presented in Table \ref{tab:04} are even more convincing in showing a significant reduction in computations as the volume of obtained search information increases. Thus, the amount of computations was reduced  139.4 times for the group, which included from 41 to 50 subproblems. \par

For one of the solved problems more complete information is presented in Table \ref{tab:05}. Row 2 contains the characteristics of the Pareto domain approximation computed numerically. In rows 3-4 the results of solving the MCO problem without the reuse of the search information are given for various values of required accuracy of computations. In the last row (row 5), the results of computations with the reuse of the search information are shown. One can see that the best results with respect to the completeness and the uniformity of approximation are presented in row 3 (the computations without the reuse of the search information). However, the number of performed iterations (the computations of the criteria values) exceeds the indicator given in row 5 more than 10 times. At approximately equal allowed number of iterations, the efficiency of computations with the reuse of the search information is higher (see rows 4 and 5). \par

In addition, an experiment on the solving two subproblems with the maximum differing values of the convolution coefficients $\lambda_1=(1,0)$ and $\lambda_2=(0,1)$ has been performed (i. e. the minimizations of the first criterion and of the second one have been performed successively). Total for finding the minimum values of the criteria without the reuse of the search information 855 iterations (the computations of the criteria values) have been executed, whereas with the reuse of the search information 530 iterations were necessary only (for the optimization of the second criterion at $\lambda_2=(0,1)$ 395 iterations have been executed in the former case, and 50 iterations only – in the latter one).\par

\begin{table}[t]
\centering
\caption{The efficiency indicators of the solving of one of subproblems from Experiment 3}
\label{tab:05}
\begin{tabular}{clcclcc}
\hline
\textbf{}                                                                                                  & \textbf{} & \multicolumn{2}{c}{\textbf{Indicators}} & \textbf{} & \textbf{\begin{tabular}[c]{@{}c@{}}Average number \\ of iterations\end{tabular}} & \multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}c@{}}Number of points\\ in PDA\end{tabular}}} \\
\textbf{}                                                                                                  & \textbf{} & \textbf{DU}        & \textbf{HV}        & \textbf{} & \textbf{per subproblems}                                                         &                                                                                             \\ \hline 
Pareto domain approx.                                                                                      &           & 0.039              & 20.852             &           &                                                                                  & 5087                                                                                        \\ \cline{1-1} \cline{3-4} \cline{6-7} 
\begin{tabular}[c]{@{}c@{}}Computations without using \\ search information ($\epsilon=0.01$)\end{tabular} &           & 0.076              & 20.847             &           & 1022.74                                                                          & 803                                                                                         \\ \cline{1-1} \cline{3-4} \cline{6-7} 
\begin{tabular}[c]{@{}c@{}}Computations without using \\ search information ($\epsilon=0.05$)\end{tabular} &           & 0.192              & 20.728             &           & 166.96                                                                           & 123                                                                                         \\ \cline{1-1} \cline{3-4} \cline{6-7} 
\begin{tabular}[c]{@{}c@{}}Computations using \\ search information ($\epsilon=0.01$)\end{tabular}         &           & 0.169              & 20.820             &           & 92.52                                                                            & 200                                                                                         \\ \hline
\end{tabular}
\end{table}

In the fourth experiment, the two-dimensional MCO problem with 5 criteria, i.e. for $N=2$, $s=5$, was solved. The problems were generated using the multiextremal functions generated by the GKLS-generator again. The following values were used: the reliability parameter $r=5$, the search accuracy $\epsilon=0.01$. As in previous experiments, MCO problems (\ref{eq:05}) were solved using 50 different convolution coefficient values. These experiments showed that the average number of optimization iterations for solving a single problem (\ref{eq:06}) without using the accumulated search information was 1,574.7. In the case when the search information was taken into account, the average number of optimization iterations executed by MAMGS was 89.2 (more than a 17-fold reduction).\par

In the final experiment, the bi-criteria five-dimensional MCO problem, i.e. for $N=5$, $s=2$, was solved. The problems were generated using the multiextremal functions generated by the GKLS-generator \cite{c14}. The following values were used: the reliability parameter $r=4$, the accuracy $\epsilon=0.03$. As in the previous experiments, MCO problems (\ref{eq:06}) were solved using 50 different convolution coefficient values. These experiments showed that the average number of optimization iterations for solving a single problem (\ref{eq:06}) without using the accumulated search information was 300,383. In the case the search information was taken into account, the average number of optimization iterations executed by MAMGS was 49,751 (more than 6-fold reduction).

\section{Conclusion}
The paper proposes an efficient method for solving complex multicriterial optimization problems, for which the optimality criteria can be multiextremal and the calculation of the criteria values can be time-consuming. The proposed approach involves the transformation of multicriterial problems to global optimization problems through the minimax convolution of the partial criteria, the reduction of dimensionality using the Peano curves, and the use of the efficient information-statistical global optimization methods.\par

The key aspect of this approach is related to overcoming the huge computational complexity of the MCO problems. A significant increase in the efficiency and a considerable reduction of the amount of computations are provided by the intensive use of all search information obtained during the process of solving multicriterial optimization problems. As a part of this approach, the methods of recalculating all available search information to the values for the subsequent global optimization problem have been proposed. The updated search information is then used in the global optimization methods for the adaptive selecting of the global search iterations. According to the results of the computational experiments, the proposed approach reduces repeatedly the computational complexity of solving the multicriterial optimization problems.\par

To conclude, it can be noted that the proposed approach is promising and requires further investigation. First of all, it is necessary to continue conducting computational experiments to address the multicriterial optimization problems with a greater number of partial criteria and for greater dimensionality in the optimization problems being solved. The potential for parallel computing should also be evaluated in view of the high computational complexity for solving global optimization problems.

\section*{Acknowledgements}  
This work has been supported by the Russian Science Foundation, project No 16-11-10150 ''Novel efficient methods and software tools for time-consuming decision making problems using superior-performance supercomputers.''

% Non-BibTeX users please use
\begin{thebibliography}{}

\bibitem{c1}	Barkalov, K., Gergel, V.: Parallel global optimization on GPU. Journal of Global Optimization, 1--18 (2015). DOI 10.1007/s10898-016-0411-y.
\bibitem{c2}	Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi Coprocessor for Solving Global Optimization Problems. LNCS, \textbf{9251}, 307--318. Springer, Heidelberg (2015)
\bibitem{c3}	Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster. Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings, \textbf{1738}, art. no. 400006 (2016)
\bibitem{c4}	Bleuler, S., Laumanns, M., Thiele, L., Zitzler, E.: PISA-A Platform and Programming Language Independent Interface for Search Algorithms. Evolutionary Multi-Criterion Optimization, LNCS, \textbf{2632}, 494--508 (2003)
\bibitem{c5}	Branke, J., Deb, K., Miettinen, K., Slowinski, R. (eds.): Multi-Objective Optimization-Interactive and Evolutionary Approaches. Springer, Berlin (2008)
\bibitem{c6}	Collette, Y., Siarry, P.: Multiobjective Optimization: Principles and Case Studies (Decision Engineering). Springer-Verlag, Berlin, Heidelberg (2011)
\bibitem{c7}	Deb, K.: Multi-Objective Optimization using Evolutionary Algorithms. Wiley, Chichester (2001)
\bibitem{c8}	Ehrgott, M.: Multicriteria Optimization. Springer-Verlag, Berlin, Heidelberg (2005, 2nd ed., 2010)
\bibitem{c9}	Eichfelder, G.: Scalarizations for adaptively solving multi-objective optimization problems. Comput. Optim. Appl., \textbf{44}, 249--273 (2009)
\bibitem{c10}	Evtushenko, Yu.G., Posypkin, M.A.: Method of non-uniform coverages to solve the multicriteria optimization problems with guaranteed accuracy. Automation and Remote Control, \textbf{75}(6), 1025--1040 (2014)
\bibitem{c11}	Figueira, J., Greco, S., Ehrgott, M. (eds.): Multiple criteria decision analysis: State of the art surveys.  Springer, New York (2005)
\bibitem{c12}	Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. Princeton University Press (2016)
\bibitem{c13}	Gablonsky, J.M., Kelley, C.T.: A Locally-Biased Form of the DIRECT Algorithm. Journal of Global Optimization, \textbf{21}(1), 27--37 (2001)
\bibitem{c14}	Gaviano, M., Lera, D., Kvasov, D.E., Sergeyev, Y.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Trans. Math. Software, \textbf{29}, 469--480 (2003)
\bibitem{c15}	Gergel, V.: A Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems. 2nd International Conference on Mathematics and Computers in Sciences and in Industry, 13--18 (2015). DOI: 10.1109/MCSI.2015.18 
\bibitem{c16}	Gergel, V., Grishagin, V., Gergel, A.: Adaptive nested optimization scheme for multidimensional global search. J. Glob. Optim., \textbf{66} (1), 35--51 (2016) 
\bibitem{c17}	Gergel, V., Lebedev, I.: Heterogeneous Parallel Computations for Solving Global Optimization Problems. Procedia Computer Science, \textbf{66}, 53--62 (2015). DOI 10.1007/s10898-016-0411-y.
\bibitem{c18}	Gergel, V., Sidorov, S.: A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems. LNCS, \textbf{9251}, 505--515. Springer-Verlag, Berlin, Heidelberg (2015)
\bibitem{c19}	Germeyer, Yu.B.: Introduction in Theory of Operation Research. Moscow, Nauka (1971) (In Russian)
\bibitem{c20}	Grishagin, V.A.: Operating characteristics of some global search algorithms. Problems of Stochastic Search, \textbf{7}, 198--206 (1978) (In Russian)
\bibitem{c21}	Grishagin, V.A., Sergeyev, Y.D., Strongin, R.G.: Parallel characteristic algorithms for solving problems of global optimization. J. Global Optim., \textbf{10}, 185--206 (1997)
\bibitem{c22}	Hill, J.D.: A Search Technique for Multimodal Surfaces. IEEE Transactions on Systems and Cybernetics, \textbf{5}(1), 2--8 (1969)
\bibitem{c23}	Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications. Surv. Math. Ind., \textbf{11}, 1--42 (2005)
\bibitem{c24}	Horst, R., Tuy, H.: Global Optimization: Deterministic Approaches. Springer-Verlag, Berlin (1990)
\bibitem{c25}	Kvasov, D.E., Pizzuti, C., Sergeyev, Y.D.: Local tuning and partition strategies for diagonal GO methods. Numerische Mathematik, \textbf{94}(1), 93-106 (2003)
\bibitem{c26}	K{\"o}ksalan, M., Wallenius, J. and Zionts, S.: Multiple criteria decision making: From early history to the 21st century. World Scientific, Singapore (2011)
\bibitem{c27}	Locatelli, M., Schoen, F.: Global Optimization: Theory, Algorithms, and Applications. SIAM (2013)
\bibitem{c28}	Mardani, A., Jusoh, A., Nor, K., Khalifah, Z., Zakwan, N., Valipour, A.: Multiple criteria decision-making techniques and their applications -- a review of the literature from 2000 to 2014. Economic Research-Ekonomska Istra\v{z}ivanja, \textbf{28}(1), 516--571 (2015). DOI: 10.1080/1331677X.2015.1075139
\bibitem{c29}	Marler, R. T., Arora, J. S.: Survey of multi-objective optimization methods for engineering. Struct. Multidisciplinary Optimization, \textbf{26}, 369--395 (2004)
\bibitem{c30}	Miettinen, K.: Nonlinear Multiobjective Optimization. Springer (1999)
\bibitem{c31}	Sergeyev, Y.D., Famularo, D., Pugliese, P.: Index Branch-and-Bound Algorithm for Lipschitz univariate global optimization with multiextremal constraints. Journal of Global Optimization, \textbf{21}(3), 317--341 (2001)
\bibitem{c32}	Sergeyev, Y.D., Grishagin, V.A.: Sequential and parallel algorithms for global optimization. Optimization Methods and Software, \textbf{3}, 111--124 (1994)
\bibitem{c33}	Sergeyev, Y.D., Grishagin, V.A.: Parallel Asynchronous Global Search and the Nested Optimization Scheme. J. Comput. Anal. Appl., \textbf{3}(2), 123--145 (2001)
\bibitem{c34}	Sergeyev, Y.D., Kvasov, D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. on Optimization, \textbf{16}(3), 910--937 (2006)
\bibitem{c35}	Sergeyev, Y.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer-Verlag, New York (2013)
\bibitem{c36}	Siwale, I.: Practical multi-objective programming. Technical Report RD-14-2013. Apex Research Limited (2014)
\bibitem{c37}	Strongin, R.G.: Numerical methods in multiextremal problems: information-statistical algorithms. Nauka, Moscow (1978) (In Russian)
\bibitem{c38}	Strongin, R.G., Sergeyev, Y.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000, 2nd ed. 2013, 3rd ed. 2014)
\bibitem{c39}	Pint{\'e}r, J.D.: Global optimization in action (continuous and Lipschitz optimization: algorithms, implementations and applications). Kluwer Academic Publishers, Dortrecht (1996)
\bibitem{c40}	Tan, K.C., Khor, E.F., Lee, T.H.: Multi-objective Evolutionary Algorithms and Applications. Springer-Verlag, London (2005)
\bibitem{c41}	T{\"o}rn, A., {\v Z}ilinskas, A.: Global Optimization. LNCS, \textbf{350}. Springer-Verlag, Berlin (1989)
\bibitem{c42}	Yang, X.-S.: Nature-inspired metaheuristic algorithms. Luniver Press, Frome (2008)
\bibitem{c43}	Zavadskas, E. K., Turskis, Z., Kildien{\.e}, S.: State of art surveys of overviews on MCDM/MADM methods. Technological and Economic Development of Economy, \textbf{20}, 165--179 (2014)
\bibitem{c44}	Zhigljavsky, A.A.: Theory of Global Random Search. Kluwer Academic Publishers, Dordrecht (1991)
\bibitem{c45}	{\v Z}ilinskas, A., {\v Z}ilinskas, J.: Adaptation of a one-step worst-case optimal univariate algorithm of bi-objective Lipschitz optimization to multidimensional problems. Commun Nonlinear Sci Numer Simulat, \textbf{21}, 89--98 (2015)
\bibitem{c46}	Zitzler, E., Knowles, J., Thiele, L.: Quality Assessment of Pareto Set Approximations. LNCS in Multiobjective Optimization-Interactive and Evolutionary Approaches. LNCS, \textbf{5252}, 373--404 (2008)
\bibitem{c47}	Strongin, R.G., Gergel, V.P. Markin, D.L.: Multicriterion multiextreme optimization with nonlinear constraints. Lecture Notes in Economics and Mathematical Systems, \textbf{351}, 120--127 (1988)
\bibitem{c48}	Chinchuluun, A., Pardalos, P.M.: A survey of recent developments in multiobjective optimization. Ann Oper Res, \textbf{154}(1), 29--50 (2007). DOI:10.1007/s10479-007-0186-0
\bibitem{c49}	Pardalos, P.M., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-Convex Multi-Objective Optimization. Springer (2017)
\bibitem{c50}	Chiandussi, G., Codegone, M., Ferrero, S., Varesio, F.E.: Comparison of multi-objective optimization methodologies for engineering applications. Computers and Mathematics with Applications, \textbf{63}(5), 912--942 (2012)
\bibitem{c51}	Zhigljavsky, A., {\v Z}ilinskas, A.: Stochastic Global Optimization. Springer (2008)
\bibitem{c52}	Paulavi{\v c}ius, R., {\v Z}ilinskas, J.: Simplicial Global Optimization. Springer (2014)

\end{thebibliography}

\end{document}
% end of file template.tex

### 2018-12-11 2016/NUMTA_MCO/NUMTA2016_kozinov_transl.tex

\documentclass{aip-cp}

\usepackage[numbers]{natbib}
\usepackage{rotating}
\usepackage{graphicx}

\usepackage{multirow}

\begin{document}

\title{Accelerating Multicriterial Optimization by the Intensive Exploitation of Accumulated Search Data}
\author[aff1]{Victor Gergel}
\eaddress{gergel@unn.ru}
\author[aff1]{Evgeny Kozinov}
\eaddress{evgeny.kozinov@itmm.unn.ru}

\affil[aff1]{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia}

\maketitle

%\textbf{Keywords.}Multicriterial decision making; global optimization problems; high performance computing; computing efficiency.

%-----------------------------------------------------------------
\begin{abstract}
The work proposed an efficient method for solving computationally difficult multicriterial optimization problems, which are widely used to model complex optimal decision making problems. Under the suggested approach, it is assumed that partial criteria can be multi-extremal and computationally intense, and finding a solution to multicriterial problems can require the sequential computation of several efficient (Pareto-optimal) alternatives. This multiple repetition of alternative searches leads to a substantial increase in computational costs, and the problem can be overcome by means of full usage of all search information obtained during the computations. The article provides a description of the developed approach, the efficiency of which has been substantiated by the results of computational experiments. 
\end{abstract}

\section{EXPANDED STATEMENT OF MULTICRITERIAL OPTIMIZATION PROBLEMS}
Multicriterial optimization (MCO) problems are used, as a rule, to model the decision making problems – see, for example, [1-5]. The opportunity to define several partial criteria complies with the decision-making practice in many areas of application. At the same time, the problem definition may need to be updated during the computations, so as to expand or reduce the number of criteria. \par

Taking into account the above, the following expanded statement of MCO problems is suggested to ensure a more adequate description of decision making problems. \par

1.	Before proceeding with the selection problem, it is necessary to define the object of optimization as the following set of components:
\begin{eqnarray} \label{eq1}
O = \langle y, D, w(y) \rangle,
\end{eqnarray}
where:
\begin{itemize}
  \item $y$ is the parameters vector:
	\begin{eqnarray*}
		y=(y_1,y_2,\dots,y_N)
	\end{eqnarray*}
		(N is the dimension of the problem);
	\item $D$ is the hyper-interval of possible values of the vector $y$, determined by the predefined vectors $a$ and $b$:	
	\begin{eqnarray*}
	   D=\{y \in R^N:a_i \leq y_i \leq b_i,1 \leq i \leq N\};
	\end{eqnarray*}
	\item $w(y)$ is the vector function of characteristics determining the object efficiently:
	\begin{eqnarray*}
		w(y)=(w_1 (y),w_2 (y),\dots,w_M (y))
	\end{eqnarray*}
\end{itemize}
(the characteristics can include, for example, weight, speed, cost, etc.). \par

It is expected that the characteristics $w_j (y)$, $1 \leq j \leq M$, can be multiextremal and require intensive computations. Without reducing the generality, the characteristics are also supposed to be non-negative, and their decreasing will result in improved object efficiency. It is also required that the model of the optimization object remains constant during the computations and does not change. \par

2. The requirements for the optimal alternative of the optimization object can be determined using the following set of elements:

\begin{eqnarray} \label{eq2}
  Z=\langle F,G,q \rangle,
\end{eqnarray}
where:
\begin{itemize}
  \item $F=<i_1,i_2,\ldots,i_s>$ is a set of indices determining the vector \textit{criteria} of efficiency
\begin{eqnarray*}
  f(y)=(w_{i_1}(y),w_{i_2}(y),\ldots,w_{i_s}(y)),
\end{eqnarray*}
as a subset of functions from the vector function of characteristics $w(y)$;
  \item $G=<j_1,j_2,\ldots,j_m>$ is a set of indices determining the vector function of \textit{constraints}  
\begin{eqnarray*}
  g(y)=(w_{j_1}(y)-q_1,w_{j_2}(y)-q_2,\ldots,w_{j_m}(y)-q_m),
\end{eqnarray*}
and $q=(q_1,q_2,\ldots,q_m )$ is the tolerance vector. The constraints $g(y)$ set the \textit{feasible search domain} 
\begin{eqnarray*}
  Q= \langle y \in D: g(y) \leq 0 \rangle .
\end{eqnarray*}
\end{itemize}

The efficiency criteria and constraints defined in this way, allow the {\it multicriterial optimization problem} to be defined
\begin{eqnarray*}
f(y) \rightarrow \min, y \in Q.
\end{eqnarray*}

This proposed scheme gives a chance to change the statement during computations. Unlike the object model (1), the MCO problem (2)  can be changed – some constraints may be modified, tolerance values may be tightened or, on the contrary, relaxed, etc. The possibility for such variations allows to set requirements for the desired decision more adequately. As a result, the decision making in the proposed approach is declared as a multi-step procedure, where the MCO problem is reviewed at each stage and subsequently solved.

\section{METHODS FOR SOLVING MCO PROBLEMS}

1. In the MCO problem (2), partial criteria are usually controversial and there are no point $y \in D$ that would provide the best (smallest) values for all criteria simultaneously. In these cases, the solution of the MCO problem considers such point $y^* \in D$, for which partial criteria cannot be improved without decreasing the efficiency for any other criteria. These unimprovable variants are called {\it efficient} or {\it Pareto optimal}. Solving an MCO problem usually requires finding one or more (in extreme cases, the entire set) of the efficient points. \par

One of the widely used approaches to finding efficient alternatives is the convolution of the partial criteria $f_i(y)$, $1 \leq i \leq s$, to a single scalar function $\Phi(\lambda, y)$, which allows to reduce the problem (2) to a scalar nonlinear programming problem (see [2-4], for example):
\begin{eqnarray}
\min \Phi(\lambda, y), y \in Q
\end{eqnarray}
where $\lambda = (\lambda_1,\lambda_2,\dots,\lambda_s)$ is the vector of coefficients used in constructing the scalar criterion. \par

A definite advantage to this approach is the possibility of solving scalar problems (3) using numerous efficient nonlinear programming algorithms [5-13]. \par

To obtain scalar criteria $\Phi(\lambda, y)$, in many cases it is possible to use a minimax convolution of the partial criteria [2-4]:
\begin{eqnarray}
\Phi(\lambda, y) = \max_{1 \leq i \leq s} \lambda_i  f_i(y),
\end{eqnarray}
\begin{eqnarray*}
\lambda_i \geq 0, 1 \leq i \leq s, \sum\limits_{i=1}^s \lambda_i = 1 
\end{eqnarray*}

A main property of minimax convolution is the necessity and sufficiency of this approach in solving the MCO problem: optimizing $F(\lambda, y)$ produces an efficient option\footnote{To be more precise, minimization of $\Phi(\lambda, y)$ can lead to  weak efficient points (the set of weak efficient points includes the Pareto domain). } for the MCO problem, and vice versa, any efficient point for the MCO problem can be obtained as a result of minimization of $F(\lambda, y)$, given the respective values of the convolution coefficients $\lambda_i, 1 \leq i \leq s$. \par
2. It should be noted that solving optimization problems (3) and (4) can require massive computations. These computations can be reduced greatly if the points for calculating criteria and constraints are only densely spaced around the globally optimal alternatives sought. This nonuniform coverage can only be supported when performing adaptive computations, when the selection of new iteration points uses the entire search data (previous iteration points and the values of the function being minimized at these points) obtained during the computations. This condition substantially complicates the computational procedures for global optimization methods, as it implies a complex computational analysis of multidimensional search data. \par

To reduce the complexity of multidimensional optimization methods, various dimensionality reduction schemes are used. For example, Peano {\it curves} or {\it mappings} $y(x)$ are widely used (see, for example, [6-7]), which allow to map uniquely the one-dimensional interval interval [0,1] onto a N-dimensional search domain. As a result of this reduction, the multidimensional global optimization problem (4) is reduced to the following univariate problem:
\begin{eqnarray}
\phi(\lambda, x) = \Phi(\lambda, y(x)), x \in [0, 1].
\end{eqnarray}

Using Peano mappings to reduce dimensionality became the grounds for developing a large number of efficient multiextremal optimization methods – see, for example, [6-12]. \par
3. Reducing dimensionality does not allow sufficient decreasing the computations in solving MCO problems. The biggest problem is that finding optimal decisions may require repeated solution of general multiextremal problems (3)-(5) when the MCO problem (2) is changed or new values for criteria convolution coefficients are selected. This problem can be overcome only by using the whole search data accumulated during the computations. \par

A numerical solution to the optimal decision making problems usually consists of repeated computations of the values of characteristics $w(y)$ at points $y^i,0 \leq i \leq k$ within the search domain $D$.The results of the respective computations can be represented as a matrix ({\it search data matrix}, SDM):
\begin{eqnarray*}
\Omega_k = \{ (y^i,w^i=w(y^i ))^T :0 \leq i \leq k \}.
\end{eqnarray*}

The size of the search data can become rather large when solving complex multidimensional problems, but any reduction in the volume of data stored, as a rule, results in excessive iterations of the global search [6-7]. \par

As a result of reducing the decision making problem to an MCO problem (2), scalarizing vector criteria (3) or (4), and reducing dimensionality (5), a SDM is transformed into a {\it search state matrix} (SSM) 
\begin{eqnarray*}
\omega_k=\{ (x_i,z_i = \phi(\lambda, x^i), f^i, g^i, w^i)^T:0 \leq i \leq k\}.
\end{eqnarray*}
which meets the statement of the optimization problem currently being solved. \par

Providing search information as an SSM is important for substantially reducing the computational intensity of solutions to the decision making problems. Optimization methods can use SSM for the adaptive performance of another round of search iterations (using the results of previous computations). And most importantly, the presence of SSM allows to recalculate the results of previous computations to the values of another next optimization problem (5) without any intense computations of the vector function characteristics $w(y)$ from (1). Thereby, all search data can be fully reused in continuing the computations. And as the main result, reusing the search data will require increasingly fewer computations for each subsequent optimization problem, until there are only a few iterations will be needed to find next efficient variants.

\section{RESULTS OF COMPUTATIONAL EXPERIMENTS}
To evaluate the efficiency of the proposed approach, a number of computational experiments for bicriterial univariate MCO problems without constraints has been executed, i.e., $N=1$, $M=2$, $s=2$, $m=0$. Multiextremal functions [14] were used as criteria for the problem defined by expression:
\begin{eqnarray*}
\psi(x) = A_0 + \sum\limits_{i=1}^{14} (A_i sin (2 i \pi x) + B_i cos (2 i \pi x)) , 0 \leq x \leq 1,
\end{eqnarray*}
were used as the problem criteria and $A_0$, $A_i$, $B_i$, $1 \leq i \leq 14$, are pseudo-random values uniformly distributedin the range of [-1,1]. \par

To optimize the convolution of the criteria (5) we used a global search algorithm developed in the framework of information-statistical theory of the global optimization [6-8], with a reliability parameter $r=2$ and search accuracy $\varepsilon=0.001$. In each experiments performed, we solved a problem (5) with several different convolution coefficients $\lambda_i$, $1 \leq i \leq s$ (from 1 to 25), with and without using the search data. The experiment results are presented in Table ~\ref{tab:tab1}.

%\begin{table}[ht]
%\centering
%\caption{Results of experiments with and without using the search data.}
%\label{tab:tab1}
%\begin{tabular}{llccccccc}
%\hline
%\multicolumn{2}{l}{Number of problems to be solved}          & 1            & 5            & 10           & 15           & 20           & 25           & 30           \\ \hline
%             Computations without  &Total number of iterations              & 40           & 234          & 566          & 746          & 1018         & 1304         & 1621         \\
%             using search data  &      Average per job        & 40           & 46.8         & 56.6         & 49.7         & 50.9         & 52.2         & 54.0         \\ \hline
%            Computations using    &       Total number of iterations       & 40           & 69           & 93           & 104          & 142          & 151          & 177          \\
%            search data   &      Average per job        & 40           & 13.8         & 9.3          & 6.9          & 7.1          & 6.0          & 5.9          \\ \hline
%\multicolumn{2}{l}{\textbf{Decreasing of executed iteration (times)}} & \textbf{1.0} & \textbf{3.4} & \textbf{6.1} & \textbf{7.2} & \textbf{7.2} & \textbf{8.6} & \textbf{9.2} \\ \hline
%\end{tabular}
%\end{table}

\begin{table}[t]
\centering
\caption{Results of experiments with and without using the search data.}
\label{tab:tab1}
\begin{tabular}{ccccccc}
\hline
  \tch{1}{c}{b}{Number of}  & \multicolumn{2}{c}{\textbf{Without using search data}}& & \multicolumn{2}{c}{\textbf{Using search data}} & \tch{1}{c}{b} {Decreasing of}  \\ \cline{2-3} \cline{5-6}  \noalign{\smallskip}
  \tch{1}{c}{b}{ problems \\to be solved}  & \tch{1}{c}{b}{Total number\\ of iterations} & \tch{1}{c}{b}{Average \\per job} & & \tch{1}{c}{b}{Total number\\ of iterations} & \tch{1}{c}{b}{Average \\per job}  & \tch{1}{c}{b} { executed \\iterations (times) }  \\ \hline
1  & 40        & 40      & & 40       & 40        & 1   \\
5  & 234       & 46,8    & & 69       & 13,8      & 3,4 \\
10 & 566       & 56,6    & & 93       & 9,3       & 6,1 \\
15 & 746       & 49,7    & & 104      & 6,9       & 7,2 \\
20 & 1018      & 50,9    & & 142      & 7,1       & 7,2 \\
25 & 1304      & 52,2    & & 151      & 6         & 8,6 \\ \hline
%30 & 1621      & 54      & & 177      & 5,9       & 9,2 \\ \hline
\end{tabular}
\end{table}

As the experiments presented above show, the use of search data allows the computational intensity to be reduced by up to 8 times when computing multicriterial optimization problems.

\section{ACKNOWLEDGEMENTS}

This research was supported by the Russian Science Foundation, project No 16-11-10150 ''Novel efficient methods and software tools for the time consuming decision making problems with using supercomputers of superior performance.''

\section{CONCLUSION}

This work proposed an approach to increasing the efficiency of handling computationally intensive multicriterial optimization problems by reusing search data obtained in the computations. The results of computational experiments confirm the viability of the developed approach. \par

Further development of this approach is needed to provide theoretical estimates for the reduction in computational intensity, and to expand substantiallythe range of computational experiments conducted for multidimensional multicriterial optimization problems with constraints.

\nocite{*}
\bibliographystyle{aipnum-cp}%
\bibliography{NUMTA2016_kozinov_transl}%

\end{document}

### 2018-12-11 2017/ICCS_MCO/ICCS_2017.tex

% easychair.tex,v 3.1 2011/12/30
%
% Select appropriate paper format in your document class as
% instructed by your conference organizers. Only withtimes
% and notimes can be used in proceedings created by EasyChair
%
% The available formats are 'letterpaper' and 'a4paper' with
% the former being the default if omitted as in the example
% below.
%
\documentclass[procedia]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

% This provides the \BibTeX macro
\usepackage{doc}
\usepackage{makeidx}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% If you plan on including some algorithm specification, we recommend
% the below package. Read more details on the custom options of the
% package documentation.
%
% \usepackage{algorithm2e}

% Some of our commands for this guide.
%
\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

\def\procediaConference{International Conference on Computational Science (12-14 June, 2017)}

%\makeindex

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Accelerating Parallel Multicriterial Optimization Methods 
Based on Intensive Using of Search Information}

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair

\titlerunning{Accelerating Parallel Multicriterial Optimization Methods}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes into the list
% defined using \institute
%
\author{
    Gergel V.P.
\and
    Kozinov E.A.
}

% Institutes for affiliations are also joined by \and,
\institute{
  Lobachevsky State University of Nizhni Novgorod\\ 
	Nizhni Novgorod, Russia \\
  \email{gergel@unn.ru, evgeny.kozinov@itmm.unn.ru} 
 }
%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{Gergel and Kozinov}

\begin{document}

\maketitle

\keywords{decision making, multicriterial optimization, parallel computing, dimensionality reduction, criteria convolution, global optimization algorithms, computational complexity}

\begin{abstract}
  In the present paper, an efficient parallel method for solving complex multicriterial optimization problems, which the optimality criteria can be multiextremal, and the computing of the criteria values can require a large amount of computations in, is proposed. The proposed approach is based on the reduction of the multicriterial problems to the global optimization ones using the minimax convolution of the partial criteria, the dimensionality reduction with the use of the Peano space-filling curves, and the application of the efficient parallel information-statistical global optimization methods. The intensive use of the search information obtained in the course of computations is provided when conducting the computations. The results of the computational experiments demonstrated such an approach to allow reducing the computation costs of solving the multicriterial optimization problems essentially -- tens and hundreds times.
\end{abstract}

%\setcounter{tocdepth}{2}
%{\small
%\tableofcontents}

%\section{To mention}
%
%Processing in EasyChair - number of pages.
%
%Examples of how EasyChair processes papers. Caveats (replacement of EC
%class, errors).

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:1}

The multicriterial optimization (MCO) problem statements are used widely for modeling the complex decision making problems -- see, for example, the monographs \cite{c5,c7,c26,c27} and the reviews of the research and application results \cite{c8,c10,c20,c25}. 
Solving the MCO problems it should be taken into account that the partial optimality criteria are usually contradictory. As a result, solving the MCO problems is reduced to finding some compromised solutions, which the achievement of the best values with respect to the partial criteria is coordinated (the compromised solutions, which cannot be improved with respect to all partial criteria are usually called {\it nondominated}, {\it efficient} or {\it Pareto-optimal} solutions). At that, the understanding of an desired compromise may change in the course of computations that leads to the necessity to find several different efficient solutions -- in the limiting case, the whole set of the nondominated solutions (the Pareto domain).\par

The necessity to find several efficient solutions (or the whole Pareto domain) increases the computational complexity essentially. One must take into account also that the partial criteria can have a complex multiextremal form, and computing the criteria values can appear to be computation costly. In such conditions, finding even a single compromised solution requires considerable computations whereas computing several efficient solutions becomes a computational problem of the ultimate level of computational complexity. Overcoming this problem becomes possible with the use of huge computational capabilities of the supercomputing systems. And, besides, to provide the efficient computations, a full utilization of the search information obtained in the course of computations is necessary. \par

Further structure of the paper is as follows. In Section \ref{sect:2}, the multicriterial optimization problem statement is given. In Section \ref{sect:3}, the basics of the developed approach are presented: the reduction of the multicriterial problems to the global optimization ones by means of the minimax convolution of the partial criteria and the dimensionality reduction with the use of the Peano space-filling curves. In Section \ref{sect:4}, a parallel multicriterial global optimization algorithm utilizing the search information accumulated in the course of computations is presented. Section \ref{sect:5} involves the results of the numerical experiments. In Conclusions, the obtained results are discussed, and main directions of further investigations are outlined.

\section{Multicriterial Optimization Problem Statement}
\label{sect:2}
A multicriterial (multi-objective) optimization (MCO) problem can be defined as follows:
\begin{eqnarray} \label{eq:01}
f(y) = (f_1(y), f_2(y),\dots, f_s(y)) \to \min,  y \in D,
\end{eqnarray}
where $y = (y_1, y_2, \dots , y_N)$ is the vector of varied parameters, $N$ is the dimensionality of the multicriterial optimization problem being solved and $D$ is the search domain representing a $N$-dimensional hyperparallelepiped 
\begin{eqnarray*}
D  = \{ y \in R^N: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}
\end{eqnarray*}
with given boundary vectors $a$ and $b$. Without loss of the generality, the partial criteria values in the problem (\ref{eq:01}) are supposed to be non-negative, and the decrease of these ones corresponds to the increase in the efficiency of the considered solutions $y \in D$.\par

As a solution of the MCO problem, any efficient solution (a {\it partial solution}) can be considered. In the general case, when solving a MCO problem, finding the whole set of nondominated solutions (the Pareto domain) $PD(f,D)$ ({\it a full solution of the MCO problem}) may be required. \par

Within the present paper, the MCO problems will be considered in application to the most complex decision making problems, which the partial criteria $f_i(y)$, $1 \leq i \leq s$ can be multiextremal, and obtaining the criteria values at the points of the search domain $y \in D$ can require a considerable amount of computations in. Let us suppose also the partial criteria $f_i(y)$, $1 \leq i \leq s$ to satisfy the Lipschitz condition
\begin{eqnarray} \label{eq:02}
|f_i(y') - f_i(y'')| \leq L_i \|y' - y''\|, y', y'' \in D, 1 \leq i \leq s. 
\end{eqnarray}

\section{Reduction Schemes for Multicriterial Problems}
\label{sect:3}
The proposed approach consists in the successive reduction of solving the MCO problems to solving a series of simpler optimization problems. In Subsection \ref{sect:3_1}, a method of the scalarization of the partial criteria, which reduces the MCO problem to a set of the global optimization ones is presented. In Subsection \ref{sect:3_2}, the dimensionality reduction with the use of the Peano space-filling curves that allows reducing the multidimensional optimization problems to the reduced one-dimensional global optimization ones is considered. \par

One can note the possibility to use a great number of efficient and widely used optimization methods for solving the MCO problems as an advantage of such an approach.

\subsection{Reducing the Multicriterial Problems to the Scalar Optimization Problems}
\label{sect:3_1}
A widely used approach for solving the MCO problems is the {\it scalarization approach}, which some methods of convolution of the set of partial criteria $f_i(y)$, $1 \leq i \leq s$ into a integrated scalar criterion $F(\lambda,y)$ are applied -- see, for example, \cite{c7,c8}. In the present paper, it is proposed to use a minimax convolution of the partial criteria in the form:
\begin{eqnarray} \label{eq:04}
F(\lambda,y) = \max_{1 \leq i \leq s}(\lambda_i f_i(y)), 
\end{eqnarray}
at given values of coefficients $\lambda_i$, $1 \leq i \leq s$, which the condition should be satisfied for
\begin{eqnarray*}
   \lambda \in  \Lambda \subset R^s:\sum_{i=1}^s{\lambda_i} = 1, \lambda_i \geq 0, 1 \leq i \leq s.
\end{eqnarray*}

The coefficients $\lambda_i$, $1 \leq i \leq s$ be understood as the indices of importance of the partial criteria -- the higher the value of the coefficient $\lambda_i$ for any particular partial criterion, the higher the contribution of this partial criterion into the integrated scalar criterion $F(\lambda,y)$. \par

It should be stressed that solving a MCO problem may require solving several problems (\ref{eq:04}) when changing the understanding of the importance of the partial criteria. The possibility to determine several efficient variants (or the whole Pareto domain) at reasonable computation costs becomes a key challenge in solving the complex multicriterial optimization problems.

\subsection{Reduction of the Multicriterial Problems to the One-dimensional Global Optimization Problems}
\label{sect:3_2}
In the general case, finding the numerical estimates of the globally-optimal extrema implies constructing a coverage of the search domain $D$ -- see, for example, \cite{c9,c11,c28,c35,c33,c42}. This fact complicates the computational schemes of the global optimization methods considerably since it implies a complex computational analysis of a large amount of multidimensional search information. As a result, many optimization algorithms utilize some methods of dimensionality reduction \cite{c13,c28,c32,c35,c33}.\par

Within the frames of the information-statistical approach, the {\it Peano curves} or {\it evolvents} $y(x)$ mapping the interval $[0,1]$ onto a $N$-dimensional hypercube $D$ unambiguously are used for the dimensionality reduction -- see, for example, \cite{c32,c35,c33}. As a result of such reduction, the initial multidimensional global optimization problem (\ref{eq:04}) is reduced to a one-dimensional one:
\begin{equation} \label{eq:05}
F(\lambda,y(x^*)) = \min_{x \in [0,1]} {F(\lambda,y(x))}.
\end{equation}

The considered dimensionality reduction scheme reduces a multidimensional problem with the Lipschitzian minimized function to a one-dimensional problem, which the corresponding functions satisfy the uniform H\"{o}lder condition (see \cite{c35,c33}) i. e.
\begin{eqnarray} \label{eq:06}
|F(\lambda,y(x')) - F(\lambda,y(x''))| \leq H |x' - x''|^{1/N}, x', x'' \in [0,1], 
\end{eqnarray}
where the constant $H$ is defined by the relation  $H = 4L \sqrt{N}$, $L$ is the Lipschitz constant from (\ref{eq:02}), and $N$ is the dimensionality of the optimization problem (\ref{eq:01}). \par

The algorithms for the numerical construction of the approximations of the Peano curves are presented in \cite{c35}.

\section{Parallel Computations for Global Optimization Problems}
\label{sect:4}
The approach presented in Section \ref{sect:3} allow reducing the solving of the MCO problem (\ref{eq:01}) to the solving a series of the reduced multiextremal problems (\ref{eq:05}). And, therefore, the problem of development of the methods for solving the MCO problems is resolved by the possibility of a wide application of the global optimization algorithms. Taking into account that the MCO problems can be computationally expensive the optimization methods should take advantage of the computational potential of high performance supercomputing systems. \par

In further description of the parallel computation methods, we will use a simpler notation of the reduced one-dimensional problems (\ref{eq:05}) as
\begin{eqnarray} \label{eq:07}
\varphi(x) = F(\lambda,y(x)): x \in [0,1]. 
\end{eqnarray}

\subsection{Parallel Algorithm of Multicriterial Global Optimization for the Computational Systems with Shared Memory}
\label{sect:4_1}
Within the framework of the proposed approach, the Parallel Multidimensional Algorithm of Global Search (PMAGS) \cite{c1,c15,c13,c16,c35,c33} is proposed to apply for solving every next reduced one-dimensional problem (\ref{eq:05}). The general computational scheme of this algorithm consists in the following.\par

Let $p$ be the number of parallel computational units (processors or cores) of a computational system with shared memory being used and let the minimized function values to be computed at any $p$ points of the interval $[0,1]$ (hereafter such computations will be called {\it trials}). Further, let $k$, $k>1$ parallel iterations of the global optimization be competed ($p$ trials are assumed to be performed within every parallel iteration). The choice of the trial points at the next $(k+1)^{th}$ parallel iteration is determined by the following rules.\par

{\it Rule 1.} Renumber the points from the set of all trial points executed earlier by the lower indices in the order of increasing coordinate value
\begin{equation} \label{eq:08}
0 = x_0 < x_1 < \dots < x_i < \dots < x_{kp} < x_{kp+1} = 1,
\end{equation}
the points $x_0$, $x_{kp+1}$ are introduced additionally for the convenience of further consideration, the minimized function values $z_0$, $z_{k+1}$ at these points remain undefined.

{\it Rule 2.} Compute the current estimate of the H\"{o}lder constant $H$ from (\ref{eq:06})
\begin{equation} \label{eq:09}
m = 
 \begin{cases}
   r M, & M > 0 \\
   1,   & M = 0
 \end{cases} ,
 M = \max_{1 \leq i \leq kp} {\frac{|z_i - z_{i-1}|} {\varrho_i} }
\end{equation}
as the relative differences of the minimized functions $\varphi(x)$ from (\ref{eq:07}) on the set of executed trial points $x_i$, $1 \leq i \leq kp$ from (\ref{eq:08}). Hereafter  $\varrho_i = (x_i - x_{i-1})^{1 / N}$, $1 \leq i \leq kp+1$. The constant $r$, $r>1$ is {\it the reliability parameter} of the algorithm.\par

{\it Rule 3.} Compute the {\it characteristics} $R(i)$ for each interval $(x_{i-1} ,x_i)$, $1 \leq i \leq kp+1$, where
\begin{equation} \label{eq:10}
\begin{matrix}
R(i) = \varrho_i + \frac{(z_i - z_{i-1})^2}{m^2 \varrho_i} - 2 \frac{(z_i + z_{i-1})}{m}, 1 \leq i \leq kp, \\
\\
R(i) = 2 \varrho_i - 4 \frac{z_{i}} {m} , i = 1, \\
\\
R(i) = 2 \varrho_i - 4 \frac{z_{i-1}} {m} , i = kp+1 \\
\end{matrix}
\end{equation}

{\it Rule 4.} Arrange the characteristics of the intervals obtained in (\ref{eq:10}) in the decreasing order 
\begin{equation} \label{eq:11}
  R(t_1) \geq R(t_2) \geq \dots \geq R(t_i) \geq \dots \geq R(t_{kp+1})
\end{equation}
and select $p$ intervals with the indices $t_j$, $1 \leq j \leq p$ having the maximal values of characteristics.\par

{\it Rule 5.} Execute new trials (the computations of the minimized function values) at the points $x^{kp+j}$ ,$1 \leq j \leq p$ located in the intervals with the maximal characteristics from (\ref{eq:11})
\begin{equation} \label{eq:12}
\begin{matrix}
x^{k+j} = \frac{x_{t_j} + x_{t_j-1}}{2} - sign(z_{t_j} - z_{t_j-1}) \frac{1}{2r} [\frac{|z_{t_j} - z_{t_j-1}|}{m}]^N, 1 < t_j < kp+1,\\
\\
x^{k+j} = \frac{x_{t_j} + x_{t_j-1}}{2}, t_j = 1 , t_j = kp + 1.
\end{matrix}
\end{equation}

The {\it termination condition}, according to which the trials are terminated, is defined by the condition 
\begin{equation} \label{eq:13}
\varrho_{t_j} < \varepsilon, 1 \leq j \leq p,
\end{equation}
which should be checked for all intervals from (\ref{eq:11}), which the scheduled trials are executed in. The parameter $\varepsilon > 0$ is the predefined {\it accuracy} of the problem solution. If the termination condition (\ref{eq:13}) is not satisfied, the iteration index $k$ is incremented by 1 (the number of the executed trials is incremented by $p$), and the next global optimization iteration is performed. \par

The characteristics $R(i)$, $1 \leq t \leq kp+1$ computed in (\ref{eq:10}) can be interpreted as some measures of importance of the intervals with respect to the location of the global minimum point in these ones. Then, the scheme of choosing the intervals for executing the next trials given by (\ref{eq:11})-(\ref{eq:12}) becomes clear -- the points for the next global optimization iteration are selected in the intervals, which the location of the global minimum point is the most probable in.\par

The conditions of convergence of the presented algorithm are considered, for example, in \cite{c35}. Thus, at an appropriate estimate of the H\"{o}lder constant ( $m > 2^{2-1/N}H$, $m$ is from (\ref{eq:09}) and $H$ is from (\ref{eq:06})) the algorithm converges to all existing global minimum points.

\subsection{Accelerating the Computations Based on Intensive Using of Search Information}
\label{sect:4_2}
Solving the multicriterial optimization problems may require a considerable volume of computations. The reason of the computational complexity consists in the fact that in order to find several efficient solutions may require multiple solving of the global optimization problems (\ref{eq:04})-(\ref{eq:05}). The overcoming of this complexity can be provided by utilization of the search information obtained in the course of computations.\par

As a result of the scalarization of the partial criteria (\ref{eq:04}), the dimensionality reduction (\ref{eq:05}), and the necessity of the ordered representation (see Rule 1 of PMAGS), the search information can be represented as a {\it search state matrix} (SSM) 
\begin{equation} \label{eq:16}
A_k=\{(x_i,z_i,v_i )^T:0 \leq i \leq kp+1\}
\end{equation} 
where $x_i$, $0 \leq i \leq kp+1$ are the reduced points of the executed global optimization iterations, $z_i$, $1 \leq i \leq kp+1$ are the values of the scalar criterion of the current optimization problem (\ref{eq:05}), and $v_i$, $1 \leq i \leq kp+1$ are the calculated values of the partial criteria $f_i(y)$, $1 \leq i \leq s$. \par

Introducing SSM causes a crucial impact on an essential reduction of the computational complexity of solving the multicriterial optimization problems. The optimization methods can use SSM for the adaptive performing the next optimization iterations taking into account the results of previous computations. And, the most important, the availability of SSM allows transforming the results of all previous computations $z_i$, $1\leq i \leq kp$ stored in SSM to the values of the current optimization problem (\ref{eq:05}) for new values of the convolution coefficients $\lambda^{'}_j$, $1 \leq j \leq s$ i.e.
\begin{equation*}
z^{'}_i=\max_{1 \leq j \leq s}⁡(\lambda^{'}_j v_i(j)),1 \leq i \leq k,
\end{equation*}
without any time-consuming computations of the partial criteria values. And, therefore, all search information can be utilized in the continued computations. \par

The PMAGS algorithm expanded with the possibility to use the search information $A_k$ from (\ref{eq:16}) will be called hereafter Parallel Multidimensional Algorithm of Multicriterial Global Search (PMAMGS).

\section{Results of Computational Experiments}
\label{sect:5}
The computational experiments have been carried out on the Lobachevsky supercomputer at Lobachevsky State University of Nizhni Novgorod (the operating system -- CentOS 6.4, the supercomputer management system -- SLURM). A computational node included 2 Intel Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM. The CPUs had 8 cores (i. e. total 16 cores were available per a node). \par

Prior to describe the computational experiments, let us consider the results of comparing the proposed approach to a number of other multicriterial optimization algorithms presented in \cite{c14}. In the comparison, the bi-criteria two-dimensional problem proposed in \cite{c9}
\begin{equation}\label{eq:17}
f_1 (y)=(y_1-1) y_2^2+1, f_2 (y)=y_2, 0 \leq y_1,y_2 \leq 1,
\end{equation}
was used. The construction of a numerical approximation $PDA(f,D)$ of the Pareto domain $PD(f,D)$ was considered as the solution of the MCO problem. In order to evaluate the quality of the approximation, the completeness and the uniformity of coverage of the Pareto domain were compared using the following two indices \cite{c9,c42}:
\begin{itemize}
	\item 	The {\it hypervolume index} (HV) defined as the volume of the subdomain of the values of the vector criterion $f(y)$ dominated by the points of the Pareto domain approximation. The value of this index can be obtained as the total volume of the intersection of the hyperparallelepipeds, the vertices of which are the points of the Pareto domain approximation, and some reference point \^{f} ̂.This index features the completeness of an approximation of the Pareto domain (the larger the value, the more complete the coverage of the Pareto domain).
	\item 	The {\it distribution uniformity index} (DU) of the points from the Pareto domain approximation defined by the relations:
\begin{equation*}
\begin{matrix}
DI=\sum_{i=1}^p{(d_i-d)^2} , d=\dfrac{ \sum_{i=1}^p{d_i} }{p}, \\
d_i=\min {\varrho(y_i,y)}, y_i,y \in DPA, y_i \neq y, 1 \leq i \leq p,
\end{matrix}
\end{equation*}
where $p$ is the number of points in the Pareto domain approximation. This index features the uniformity of a coverage of the Pareto domain (the less the value, the more uniform the coverage of the Pareto domain).

\end{itemize}
Within the frames of this experiment, five multicriterial optimization algorithms have been compared:
\begin{itemize}
\item The Monte-Carlo (MC) method, in which the trial points were chosen randomly and uniformly within the search domain $D$,
\item The genetic algorithm SEMO from the PISA library \cite{c20,c42},
\item The Non-Uniform Coverage (NUC) method \cite{c20},
\item The Bi-objective Lipschitz Optimization (BLO) method proposed in \cite{c42},
\item The MAMGS (the serial version of PMAMGS) algorithm proposed in the present paper.
\end{itemize}
For the first three algorithms from the above list, the results of experiments from \cite{c20} have been used. For the BLO method, the results of experiments were presented in \cite{c42}. For MAMGS the following parameters were used: the reliability parameter $r = 3$ and the accuracy  $\varepsilon = 0.05$. Total 50 problems (\ref{eq:05}) with various convolution coefficients $\lambda$ distributed uniformly in $\Lambda$ from (\ref{eq:04}) have been solved by MAMGS.\par

The results of experiments from \cite{c14} are presented in Table \ref{tab:01}. The results have demonstrated that the MAMGS algorithm has a considerable advantage as compared to considered multicriterial optimization methods even when solving relatively simple MCO problems.

\begin{table}[t]
\centering
\caption{The results of computational experiments from \cite{c14} for the test problem (\ref{eq:17})}
\label{tab:01}
\begin{tabular}{cllll}
\hline
\textbf{Method} & \textbf{Iterations} & \textbf{PDA points} & \textbf{HV} & \textbf{DU} \\ \hline
MC              & 500                 & 67                  & 0.300       & 1.277       \\
SEMO            & 500                 & 104                 & 0.312       & 1.116       \\
NUC             & 515                 & 29                  & 0.306       & 0.210       \\
BLO             & 498                 & 68                  & 0.308       & 0.175       \\
MAMGS+          & 370                 & 100                 & 0.316       & 0.101       \\ \hline
\end{tabular}
\end{table}

In the executed series of experiments, the solving of bi-criteria two-dimensional MCO problems i.e. $N=2$, $s=2$ has been performed. The multiextremal functions defined below:
\begin{equation*}
\begin{matrix}
AB = \Big(\sum_{i=1}^7 \sum_{j=1}^7 [A_{ij} a_{ij}(y_1, y_2) + B_{ij} b_{ij}(y_1, y_2)]\Big)^2 \\
CD = \Big(\sum_{i=1}^7 \sum_{j=1}^7 [C_{ij} a_{ij}(y_1, y_2) - D_{ij} b_{ij}(y_1, y_2)]\Big)^2 \\ 
\phi(y_1, y_2) = - (AB + CD)^{1/2}
\end{matrix}
\end{equation*}
were used as the problem criteria. Here 
\begin{equation*}
\begin{matrix}
a_{ij}(y_1, y_2) = \sin(\pi i y_1) \sin(\pi j y_2), &
b_{ij}(y_1, y_2) = \cos(\pi i y_1) \cos(\pi j y_2)
\end{matrix}
\end{equation*}
are defined in the range $0 \leq y_1, y_2 \leq 1$, and the parameters  $-1\leq A_{ij},B_{ij},C_{ij},D_{ij} \leq 1$ are independent values distributed uniformly and randomly. The minimization of such functions arises, for example, in the problem of evaluation of the maximum strain (determined by the strength) in a thin plate at the transverse load. \par

In order to provide more justified conclusions on the efficiency of the developed approach, the solving of 100 multicriterial problems formed with the use of the multiextremal functions of the above problem family has been performed. For solving the optimization problems, the reliability parameter $r=2$ and the search accuracy $\varepsilon=0.001$ were used. In every experiment performed, the problem (\ref{eq:04}) has been solved for 50 various convolution coefficients with the use of the search information and without the one. The obtained results of experiments were averaged over the number of solved MCO problems (i. e. the results of experiments are presented below in average per one MCO problem, which the global search of the Pareto optimal solutions for 50 problems (\ref{eq:04}) has been performed within). As an example, an approximation of the Pareto domain obtained for a problem to be solved is shown in Figure \ref{fig:F1F2_gr}.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{F1F2_gr}
	\caption{Approximation of the Pareto domain for a MCO problem to be solved}
	\label{fig:F1F2_gr}
\end{figure}

The results of numerical experiments are presented in Table \ref{tab:02}.

\begin{table}[t]
\centering
\caption{The results of the series of experiments on solving the bi-criteria two-dimensional MCO problems}
\label{tab:02}
\begin{tabular}{cccccccc}
\hline
\multicolumn{1}{c}{\textbf{}}  &  & \multicolumn{2}{c}{\textbf{Iterations}}                                  &  & \multicolumn{3}{c}{\textbf{Speedup}}                                                                                   \\ \cline{3-4} \cline{6-8} 
\multicolumn{1}{c}{\textbf{p}} &  & \multicolumn{1}{c}{\textbf{PMAGS}} & \multicolumn{1}{c}{\textbf{PMAMGS}} &  & \multicolumn{1}{c}{\textbf{PMAGS}} & \multicolumn{1}{c}{\textbf{PMAMGS - 1}} & \multicolumn{1}{c}{\textbf{PMAMGS - 2}} \\ \cline{1-1} \cline{3-4} \cline{6-8}
1                              &  & 2                                  & 3                                   &  & 4                                  & 5                                       & 6                                       \\ \cline{1-1} \cline{3-4} \cline{6-8}
1                              &  & 17161,9                            & 1773,4                              &  & 1                                  & 1                                       & 9,7                                     \\
2                              &  & 9019,4                             & 1103,1                              &  & 1,9                                & 1,6                                     & 15,6                                    \\
5                              &  & 4117,2                             & 550,1                               &  & 4,2                                & 3,2                                     & 31,2                                    \\
10                             &  & 1801,7                             & 179,6                               &  & 9,5                                & 9,9                                     & 95,6                                    \\
25                             &  & 820,2                              & 77,5                                &  & 20,9                               & 22,9                                    & 221,4                                   \\ \hline
\end{tabular}
\end{table}

To explain the presented results, let us remind that PMAGS is a parallel algorithm without the use of the search information while PMAMGS uses the search information intensively. \par

The results of experiments demonstrate an essential reduction of the amount of computations by means of the use of the search information obtained in the course of the global optimization. Even for the serial algorithm without any use of the parallel computations, this reduction reaches more than 9 times (Column 6). In order to account for this result, two variants of obtained speedup are given in Table \ref{tab:02}: one relative to the serial algorithm utilizing the search information (Speedup 1, Column 5) and another one relative to the initial serial algorithm (Speedup 2, Column 6). As follows from the presented data, the speedup of the parallel computations with the use of the search information was 22.9 when using 25 computational cores. One should note a considerable increase of the speedup of the parallel computations relative to the initial algorithm without the use of the search information -- the speedup was 221.4 for 25 computational cores. In addition, the obtained results are presented also in the visual graphical form in Figure \ref{fig:res_exp}.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{res_exp}
	\caption{Diagrams for demonstrating the results of the numerical experiments}
	\label{fig:res_exp}
\end{figure}

In Figure \ref{fig:res_exp}, the bar chart demonstrates the number of parallel iterations executed by the methods (columns 2 and 3 in Table 2), the corresponding numbers are shown on the left axis. The line plots present the speedups of parallel computations (columns 5 and 6 in Table 2), the corresponding scale is shown on the right axis.

\section{Conclusions}
\label{sect:6}
In the present paper, an efficient parallel method for solving complex multicriterial optimization problems, which the criteria of optimality can be multiextremal, and the computing of the criteria values can require a large volume of computations in, has been proposed. The basis of the proposed approach consists in reduction of the multicriterial problems to the global optimization ones by means of the minimax convolution of the partial criteria, the dimensionality reduction with the use of the Peano space-filling curves, and the application of the efficient information-statistical global optimization methods. \par

The key aspect of the developed approach consists in the overcoming of large computations costs of the global optimization of the set of efficient solutions when solving the multicriterial optimization problems. A considerable increase of the efficiency and essential reduction of the amount of computations are provided by means of intensive using of the search information obtained in the course of computations. Within the framework of developed approach, a method for transforming all available search information to the values of current scalar problem of global optimization being solved have been proposed. The search information is used by the parallel optimization methods for the adaptive planning of the performed global optimization iterations. The results of the computational experiments demonstrated such an approach to allow reducing the computational costs of solving the multicriterial optimization problems essentially -- by tens and hundreds times. \par

In conclusion, one can note that the developed approach is a promising one and requires further investigations. First of all, it is necessary to continue carrying out the computational experiments on solving the multicriterial optimization problems with a large number of partial criteria of efficiency and for a greater dimensionality of the optimization problems being solved. It is necessary also to estimate the possibility to organize the parallel computations with the use of a greater number (hundreds and thousands) of processors in the high performance computational systems with the distributed memory.

\section{Acknowledgements}
\label{sect:7}

This work has been supported by Russian Science Foundation, project No 16-11-10150 ''Novel efficient methods and software tools for time-consuming decision making problems using superior-performance supercomputers.''
%------------------------------------------------------------------------------
% Refs:
%
\label{sect:bib}
\bibliographystyle{plain}
%\bibliographystyle{alpha}
%\bibliographystyle{unsrt}
%\bibliographystyle{abbrv}
\bibliography{iccs_2017}

%------------------------------------------------------------------------------
\end{document}

% EOF

### 2018-12-11 2018/RSCD_MCO/RSCD_MCO_2018.tex

\documentclass[runningheads]{llncs}

% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{bbding}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{GPU-based Parallel Computations in Multicriterial Optimization 
%\thanks{This research was supported by the Russian Science Foundation, project No 16-11-10150 ''Novel efficient methods and software tools for time-consuming decision-making problems using supercomputers of superior performance.''}
}

\author{ Victor Gergel \orcidID{0000-0002-4013-2329}(\Envelope) \and
Evgeny Kozinov \orcidID{0000-0001-6776-0096}}

\authorrunning{V. Gergel et al.}

\institute{Lobachevsky State University of Nizhni Novgorod \\
Nizhni Novgorod, Russia \\
\email{gergel@unn.ru}, \email{evgeny.kozinov@itmm.unn.ru}}

\maketitle

\begin{abstract}
In the present paper, an efficient approach for solving the time-consuming multicriterial optimization problems, in which the optimality criteria could be the multiextremal ones and computing the criteria values could require a large amount of computations is proposed. The proposed approach is based on the reduction of the multicriterial problems to the scalar optimization ones with the use of the minimax convolution of the partial criteria, on the dimensionality reduction with the use of the Peano space-filling curves, and on the application of the efficient information-statistical global optimization methods. An additional application of the block multistep scheme provides the opportunity of the large-scale parallel computations with the use of the graphics processing units (GPUs) with thousands of computational cores. The results of the numerical experiments have demonstrated such an approach to allow improving the computational efficiency of solving the multicriterial optimization problems considerably -- hundreds and thousands.
\end{abstract}

\keywords{Decision making \and Multicriterial optimization \and Global optimization \and High performance computations \and Dimensionality reduction \and Criteria convolution \and Global search algorithm \and Computational costs}

\section{Introduction} \label{sec:01}

The multicriterial optimization (MCO) problems are classified as the most general statements of the decision-making problems -- the statement of MCO problems covers many classes of optimization problems, including unconstrained optimization, nonlinear programming, global optimization, etc. The opportunity to set several criteria is very useful in the formulating of the complex decision-making problems and is used in the applications widely. Such practical importance has caused a high activity of research in the field of the MCO problems. As a result of the performed investigations, a large number of the efficient methods of solving the MCO problems have been proposed and a great number of the applied problems have been solved -- see, for example, the monographs \cite{c2,c3,c4,c5} and the reviews of the scientific and practical results in the field \cite{c7,c9}.

The present paper is devoted to the solving of the MCO problems, which are used for formulating the decision-making problems in the design of complex technical objects and systems. In such applications, the partial criteria could take a complex multiextremal form, and computing the values of the criteria could require a large amount of computations. In such conditions, finding even a single efficient decision requires a significant amount of computations whereas finding several decisions (or the complete set of these ones) becomes a problem of high computation costs.

Among the directions used for solving the MCO problems widely, the scalarization approach utilizing some methods of the partial criteria convolution to a single scalar criterion is applied -- see, for example, \cite{c3,c5,c10}. Among such approaches, there are the methods of finding the decisions, which are the closest to the ideal one or to the compromised ones, or to the prototypes existing actually, etc. Among such algorithms, there exists the method of successive concessions, in which some tolerances to possible values of criteria are introduced. The scalarization of a vector criterion allows reducing the solving of a MCO problem to solving a series of the multiextremal optimization problems and, therefore, utilizing all existing highly efficient global search algorithms for the multicriterial optimization.

One of the promising approaches to solving the time-consuming global optimization problems consists in utilizing the graphics processing units (GPUs). At present, a GPU is a high-performance flexible programmable massive parallel processor, which can provide solving many complex computational problems \cite{c18}. However, the use of the GPU computational potential in the field of global optimization is quite limited. As a rule, GPUs are used for the parallelization of the algorithms, which are based on the random search concept in any way (see \cite{c19,c20,c21}). A review of this direction is given in \cite{c22}.

Further structure of the paper is as follows. In Section \ref{sec:02}, the multicriterial optimization problem statement is given and the basics of the developed approach are considered, namely the reduction of the multicriterial problems to the scalar optimization ones using the minimax convolution of the partial criteria and the dimensionality reduction using the Peano space-filling curves. In Section \ref{sec:03}, the parallel global search algorithm for solving the reduced scalar optimization problems is described and the block multistep scheme of the dimensionality reduction, which provides the opportunity to use the GPUs with thousands of computational cores is presented. Section \ref{sec:04} includes the results of numerical experiments confirming the proposed approach to be a promising one. In Conclusion, the obtained results are discussed and main possible directions of further investigations are outlined.

\section{Problem Statement}\label{sec:02}

The multicriterial  optimization (MCO) can be defined as follows:
\begin{equation}
\label{eq:01}
f(y) = (f_1(y), f_2(y),\dots, f_s(y)) \to min,  y \in D,	
\end{equation}
where $y = (y_1, y_2,\dots, y_N)$ is the vector of the varied parameters, $N$ is the dimensionality of the multicriterial optimization problem being solved, $f(y)$ is the vector efficiency criterion, and $D$ is the search domain representing an $N$-dimensional hyperparallelepiped 
\begin{equation}
D  = { y \in R^N: a_i \leq y_i \leq b_i, 1 \leq i \leq N }
\end{equation}
at given boundary vectors $a$ and $b$.

Without loss of generality, the values of partial criteria in the problem (\ref{eq:01}) are supposed to be non-negative, and the decreasing of these ones corresponds to increasing efficiency of the decisions $y \in D$.

In the present work the problem (\ref{eq:01}) will be considered in application to the most complex decision-making problems, in which the partial criteria $f_i(y)$, $1 \leq i \leq s$ could be multiextremal, and obtaining the criteria values at the points of the search domain $y \in D$ could require a considerable amount of computations. Let us suppose also the partial criteria $f_i(y)$ to satisfy the Lipschitz condition
\begin{eqnarray} \label{eq:02}
|f_i(y') - f_i(y'')| \leq L_i \|y' - y''\|, y', y'' \in D, 1 \leq i \leq s. 
\end{eqnarray} 
where $L_i$ is the Lipschitz constant for the functions $f_i(y)$, $1 \leq i \leq s$ and $\| * \|$ denotes the Euclidean norm in $R^N$.

The general approach to solving the MCO problem applied in the present work consists in the reduction of solving the MCO problems to the solving of a series of one-dimensional optimization problems: 
\begin{eqnarray} \label{eq:03}
\min{⁡ \phi(x)=F(\lambda, y(x)),x \in [0,1]},
\end{eqnarray} 
where 
\begin{eqnarray}
F(\lambda, y(x)) = \max{((\lambda_i f_i (y(x)), 1 \leq i \leq s)}
\end{eqnarray} 
is the minimax convolution of the partial criteria of the MCO problem with the use of the vector of the convolution coefficients 
\begin{eqnarray}
\lambda=(\lambda_1,\lambda_2,\dots,\lambda_s) \in \Lambda \subset R^s : 
\sum_{i=1}^s{\lambda_i=1}, \lambda_i \geq 0, 1 \leq i \leq s	
\end{eqnarray}
and $y(x)$ is a continuous and unambiguous mapping of the interval $[0,1]$ onto the $N$-dimensional search domain $D$ -- see, for example, \cite{c11,c12,c13}.

\section{GPU-based Parallel Computations for Solving the Multicriterial Optimization Problems}\label{sec:03}

The convolution of the partial criteria applied within the framework of the developed approach and the dimensionality reduction allow reducing the solving of the MCO problem (\ref{eq:01}) to solving a series of the reduced multiextremal problems (\ref{eq:03}). And, therefore, the problem of the development of the methods for solving the MCO problems is resolved by the opportunity of a wide use of the global search algorithms. The state of the art in the field of global optimization is presented comprehensively enough, for example, in \cite{c11,c27,c28}.

In the present work, the global search algorithm developed within the framework of the information-statistical theory of the multiextremal optimization is proposed to use for solving the reduced problems (\ref{eq:03}). This theory served as a basis for the development of a large number of algorithms, which have been substantiated mathematically, have demonstrated high efficiency, and allowed solving many complex optimization problems in various fields of application \cite{c11,c12,c29,c30,c32,c33}.

The approach proposed for the organization of the parallel computations when solving the time-consuming multiextremal optimization problems is based on the simultaneous computing of the partial criteria values in the MCO problem (\ref{eq:01}) at several different points of the search domain $D$. Such an approach provides the parallelization of the most time-consuming part of the global search and is a general one -- it can be applied for many global search methods in various global optimization problems.

\subsection{Parallel Algorithm of Global Search for Finding the Efficient Decisions in the Multicriterial Optimization Problems}

Within the framework of this approach, the multidimensional generalized parallel algorithm of global search (PAGS) for finding the efficient decisions of the multicriterial optimization problems constitutes the basis for the developed optimization methods. The general computational scheme of the algorithm can be described as follows \cite{c11,c12,c13}.

Let $p$ is the number of employed parallel computers (processors or cores) of a computational system with shared memory. The initial two iterations of the algorithm are performed at the boundaries of the interval $x^0=0$, $x^1=1$. Besides these boundary points, the algorithm should perform additional iterations at the points $x^i$, $1<i \leq p$, which can be defined \textit{a priori} or computed by any auxiliary computational procedure. Then, let $k$, $k>p$ global search iterations have been completed, at each of which the computing of the value of the minimized function $\phi(x)$ from (\ref{eq:03}) (hereafter called \textit{a trial}) has been performed. The choice of the points for trials performed within the next iteration in parallel is determined by the following rules.

\textit{Rule 1.} Renumber the trials points of the completed search iterations by the lower indices in the order of increasing coordinate values
\begin{eqnarray} \label{eq:05}
	0=x_0<x_1<\dots<x_i<\dots<x_k=1.
\end{eqnarray}

\textit{Rule 2.} Compute current estimate of the H\"{o}lder constant of the reduced function $\phi(x)$:
\begin{eqnarray} \label{eq:06}
	m = 
 \begin{cases}
   r M, & M > 0 \\
   1,   & M = 0
 \end{cases} ,
 M = \max_{1 \leq i \leq k} {\frac{|z_i - z_{i-1}|} {\varrho_i} }
\end{eqnarray}
where $z_i=\phi(x_i )$, $\varrho_i=\sqrt[N]{x_i-x_{i-1}}$, $1 \leq i \leq k$. The constant $r$, $r>1$ is the \textit{reliability parameter} of the algorithm.

\textit{Rule 3.} Compute \textit{the characteristic} $R(i)$ for each interval $(x_{i-1}  ,x_i)$, $1 \leq i \leq k$ according to the expression 
\begin{eqnarray} \label{eq:07}
R(i) = \varrho_i + \frac{(z_i - z_{i-1})^2}{m^2 \varrho_i} - 2 \frac{(z_i + z_{i-1})}{m}, 1 \leq i \leq k,
\end{eqnarray}

\textit{Rule 4.} Arrange the characteristics of the intervals $(x_{i-1},x_i)$, $1 \leq i \leq k$ obtained according to (\ref{eq:07}) in the decreasing order 
\begin{eqnarray} \label{eq:08}
R(t_1) \geq R(t_2) \geq \dots \geq R(t_{k-1}) \geq R(t_k)
\end{eqnarray}
and select $p$ intervals with the indices $t_j$, $1 \leq j \leq p$ having the maximum values of the characteristics.

\textit{Rule 5.} Perform new trials at the points $x^{k+j}$, $1 \leq j \leq p$ placed in the intervals with the maximum characteristics from (\ref{eq:08}) according to the expressions

\begin{eqnarray} \label{eq:09}
	x^{k+j} = \frac{x_{t_j} + x_{t_j-1}}{2}
	- sign(z_{t_j} - z_{t_j-1}) \frac{[\frac{|z_{t_j} - z_{t_j-1}|}{m}]^N}{2r} , 1\leq j \leq p.
\end{eqnarray}

\textit{Stopping condition} for the algorithm, according to which the execution of the algorithm is terminated, consists in checking the lengths of the intervals, in which the scheduled trials are performed, with respect to the required \textit{accuracy} of the problem solution i.e.
\begin{eqnarray} \label{eq:10}
\varrho_t \leq \varepsilon, 1 \leq t_j \leq p.
\end{eqnarray}

Various modifications of this algorithm and the corresponding theory of convergence are presented in \cite{c11,c12}.

\subsection{Multilevel Decomposition of the Parallel Computations}

Further development of the methods of the parallel computations in the multicriterial global optimization problems and, therefore, the expansion of the possible quantity of the employed processors/cores can be ensured by the use of one more dimensionality reduction method in the decomposition scheme of the MCO problems (\ref{eq:01}) -- the multistep scheme of decomposition of the optimization problems \cite{c11,c12,c33,c34}. According to this scheme, the solution of a multidimensional optimization problem can be obtained by solving a series of «nested» one-dimensional problems:

\begin{eqnarray} \label{eq:11}
\min{\{\phi(y): y \in D\}} = 
\min_{a_1 \leq y_1 \leq b_1}{ \min_{a_2 \leq y_2 \leq b_2}{ \dots \min_{a_N \leq y_N \leq b_N}{ \phi(y) } } }.
\end{eqnarray}

The original multistep reduction scheme (\ref{eq:11}) can be generalized for the use in combination with the dimensionality reduction scheme based on the Peano curves \cite{c36}. According to the generalized block multistep scheme, the vector of variables $y \in D$ of the global optimization problem (\ref{eq:01}) is considered as a set of the block variables 
\begin{eqnarray} 
y = (y_1, y_2, \dots, y_N) = (u_1, u_2, \dots, u_M),
\end{eqnarray}
where the $i^{th}$ block variable $u_i$ is a vector with the dimensionality $N_i$ of the elements of the vector $y$ taken sequentially i.e. 
\begin{eqnarray}
\begin{aligned}
u_{i}=(y_{n_i+1},y_{n_i+2},\dots,y_{n_i+N_{i}}), \\
n_0=N_0=0, n_i=n_{i-1}+N_{i-1} ,1 \leq i \leq M
\end{aligned}
\end{eqnarray}
and $N_1 + N_2 + \dots + N_M = N$.

Using the new variables, main equation of the multistep reduction scheme (\ref{eq:11}) can be rewritten in the form
\begin{eqnarray} \label{eq:12}
\min{\{\phi(y): y \in D\}} = 
\min_{u_1 \in D_1}{ \min_{u_2 \in D_2}{ \dots \min_{u_M \in D_M}{ \phi(y) } } }.
\end{eqnarray}
where the subdomains $D_i$, $1 \leq i \leq M$ are the projections of the initial search domain $D$ onto the subspace corresponding the variables $u_i$, $1 \leq i \leq M$. As a result, in the generalized block multistep reduction scheme, the nested subproblems
\begin{eqnarray} \label{eq:13}
\phi_i(u_1 \dots u_i) = 
\min_{u_{i + 1} \in D_{i + 1}}{\phi_{i+1}(u_1 \dots u_i, u_{i+1})},1 \leq i \leq M-1
\end{eqnarray}
are the multidimensional ones, and  the dimensionality reduction method based on the Peano curves can be applied to solve these ones.

To provide the parallel computations in the block multistep reduction scheme, at each decomposition level one can generate several optimization problems simultaneously for the parallel solving of these ones \cite{c34,c36} (see Figure \ref{fig:01}). The resulting set of problems to be solved in parallel can be controlled by means of the predefined parallelization vector
\begin{eqnarray} \label{eq:14}
\pi = (\pi_1, \pi_2, \dots, \pi_M),
\end{eqnarray}
where $\pi_i$, $1 \leq i < M$ is the number of subproblems being solved in parallel at the $(i+1)^{th}$ level of decomposition arising as a result of performing the parallel iterations at the $i^{th}$ level. For the $M^{th}$ level, the quantity $\pi_M$ means the number of parallel trials in the course of minimization of the function 
\begin{eqnarray}
\phi_M(u_1, \dots, u_M) = \phi(y_1, \dots, y_N)
\end{eqnarray}
with respect to the variable $u_M$ at fixed values $u_1, \dots, u_{M-1}$, i.e. the number of values of the objective function $\phi(y)$ computed in parallel. Then, the total number of the employed processors/cores will be 
\begin{eqnarray} \label{eq:15}
\prod{} = 1 + \sum_{i=1}^{M-1}\prod_{j=1}^i{\pi_j}
\end{eqnarray}
The resulting multilevel scheme of parallel computations allows ensuring the efficient employment of all processors/nodes available in the high-performance systems with a large number of the computational nodes (including the ones with the distributed memory). The generation of a large number of optimization problems solved in parallel initiates a promising direction on a wide employment of GPUs with a large number of computational cores. This direction has been tested in solving the time-consuming global optimization problems \cite{c38,c39,c40}. In the present work, the possibility of utilizing the GPUs for solving the multicriterial optimization problems has been evaluated.

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.5\textwidth]{fig1.png}}
\caption{General scheme of generating the parallel problems using the block multistep dimensionality reduction scheme}
\label{fig:01}
\end{figure}

The PAGS algorithm combined with the block multistep scheme of dimensionality reduction will be called hereafter Generalized Parallel Algorithm of Global Search (GPAGS).

\section{Results of Numerical Experiments}\label{sec:04}

The numerical experiments have been carried out on the «Lobachevsky» supercomputer at State University of Nizhni Novgorod (operating system -- CentOS 6.4, management system -- SLURM). Each supercomputer node included 2 Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM processors. The central processor units were the 8-core one (i. e. total 16 CPU cores per a node were available). At each node, three NVIDIA Kepler K20Х GPUs were installed. To provide parallel computations MPI and CUDA technologies are applied.

The evaluation of the efficiency of the developed approach for solving the MCO problems without using the computational accelerators has already been performed earlier \cite{c14,c15,c16,c17}. Let us consider, for example, the results of experiments from \cite{c16}. For comparison, a bicriterial test problem proposed in \cite{c41} was used:

\begin{eqnarray} \label{eq:16}
f_1 (y)=(y_1-1) y_2^2+1, f_2 (y)=y_2, 0 \leq y_1,y_2 \leq 1.
\end{eqnarray}

As a solution of this MCO problem, the construction of a numerical approximation of the Pareto domain (PDA) was considered. For evaluating the quality of approximation, the completeness and the uniformity of the coverage of the Pareto domain were compared with the use of the following two indicators \cite{c16,c41}:

\begin{itemize}
\item The \textit{hypervolume index} (HV) defined as the volume of the subdomain of the values of the vector criterion $f(y)$ dominated by the points of the Pareto domain approximation. This indicator characterizes the completeness of the Pareto domain approximation (a higher value corresponds to more complete coverage of the Pareto domain).
\item The \textit{distribution uniformity} index (DU) of the points from the Pareto domain approximation. This indicator characterizes the uniformity of coverage of the Pareto domain (a lower value corresponds to more uniform coverage of the Pareto domain).
\end{itemize}

Within the described experiment, five multicriterial optimization algorithms were compared: the Monte-Carlo (MC) method, the genetic algorithm SEMO from the PISA library \cite{c7,c42}, the Non-uniform coverage (NUC) method \cite{c7}, the bi-objective Lipschitz optimization (BLO) method proposed in \cite{c42}, and the serial variant of the GPAGS algorithm proposed in the present paper.

Total for the GPAGS algorithm 50 subproblems were solved at various values of the convolution coefficients $\lambda$ distributed in $\Lambda$  uniformly. The results of experiments from \cite{c16} are presented in Table \ref{tab:01}.

\begin{table}[htbp]
\centering
\caption{Results of numerical experiments from \cite{c16} for the test problem (\ref{eq:16})}
\label{tab:01}
\begin{tabular}{ccccc}
\hline
\textbf{Method} & \textbf{Iterations} & \textbf{PDA points} & \textbf{HV} & \textbf{DU} \\ \hline
MC              & 500                 & 67                  & 0.300       & 1.277       \\
SEMO            & 500                 & 104                 & 0.312       & 1.116       \\
NUC             & 515                 & 29                  & 0.306       & 0.210       \\
BLO             & 498                 & 68                  & 0.308       & 0.175       \\
GPAGS           & 370                 & 100                 & 0.316       & 0.101       \\ \hline
\end{tabular}
\end{table}

The results of performed experiments have demonstrated the GPAGS algorithm to have a considerable advantage as compared to considered multicriterial optimization methods even in solving relatively simple MCO problems.

In the present paper, the numerical experiments were conducted in order to evaluate the efficiency of the developed approach in solving the MCO problems with the use of the GPUs. In the conducted series of experiments, the solving of the bicriterial six-dimensional MCO problems (i.e. $N = 6$, $s = 2$) has been performed. As the test problem criteria, the multiextremal functions obtained with the use of the GKLS generator \cite{c43} were used.

In the course of experiments, 50 multicriterial problems of this class have been solved. In each problem, the search of the Pareto-optimal decisions was performed for 10 convolution coefficients $\lambda$ from (\ref{eq:03}) distributed in $\Lambda$ uniformly (i.e. 500 global optimization subproblems have been solved). In solving the problems, two levels of the block dimensionality reduction scheme were used. At the first level of the reduction scheme, the optimization with respect to the first two variables was performed (i.e. $u_1 = (y_1, y_2)$, $N_1=2$). The optimization with respect to the rest variables was performed at the second decomposition level (i.e. $u_2 = (y_3, y_4, y_5, y_6)$, $N_2=4$). Trials at the second decomposition level are executed on GPU. Computations are implemented in accordance with the master/slave scheme which is not required any data communication between GPUs. Trial points are sent by CPU just before every iteration and are stored in the GPU global memory.

As the parameters the accuracy of the method $\varepsilon=0.025$ and the reliability of the method $r=6.5$ for the first level of the block reduction scheme and $r=4.5$ for the second level of decomposition were used. The results of the numerical experiments are presented in Table \ref{tab:02}.

\begin{table}[htbp]
\centering
\caption{Comparison of the times of solving of the six-dimensional bicriterial MCO problems}
\label{tab:02}
\begin{tabular}{ccccccc}
\hline
\textbf{Nodes} & \textbf{P} & \textbf{Th} & \textbf{P*Th} & \textbf{Core type} & \textbf{Time, s} & \textbf{Speedup} \\ \hline
1              & 1          & 16          & 16            & CPU                   & 7 186.4          & 1.0              \\
16             & 16         & 16          & 256           & CPU                   & 957.3            & 7.5              \\
16             & 32         & 4 032       & 129 024       & GPU                   & 529.9            & 13.6             \\
16             & 64         & 2 016       & 129 024       & GPU                   & 291.8            & 24.6             \\
16             & 128        & 1 008       & 129 024       & GPU                   & 272.4            & 26.4             \\
32             & 128        & 2 016       & 258 048       & GPU                   & 214.9            & 33.4             \\
32             & 256        & 1 008       & 258 048       & GPU                   & 253.2            & 28.4             \\ \hline
\end{tabular}
\end{table}

In Table \ref{tab:02}, the column ''Nodes'' shows the number of the supercomputer nodes employed, the column $P$ shows the number of the parallel subproblems generated at the first level of the block reduction scheme (the parameter $\pi_1$ from (\ref{eq:14})), the column $Th$ shows the number of generated points of trials performed in parallel at the second decomposition level (the parameter $\pi_2$ from (\ref{eq:14})).

The results of experiments presented in the first row show the averaged time of solving of the MCO problems with the use a single computational node of the supercomputer ($\pi_1=1$, $\pi_2=16$). In the second row of the table, the averaged time of solving of the MCO problems with the use of sixteen computational nodes of the supercomputer ($\pi_1=16$, $\pi_2=16$) is given. As follows from the presented results, the resulting speedup of the computations was 7.5 times. In the rows 3-7, the averaged times of solving the MCO problems with the use of 16 and 32 supercomputer nodes are given. At each node, 3 GPUs were employed. The number of parallel subproblems generated at the first level of the block reduction scheme (the parameter $\pi_1$ from (\ref{eq:14})) ranged from 32 to 256 whereas the number of generated points of trials performed in parallel at the second decomposition level (the parameter $\pi_2$ from (\ref{eq:14})) ranged from 1008 to 4032. Total number of employed GPU cores was 129024 with using 16 nodes, and 258048 with using 32 nodes. The maximal speedup of computations achieved was 33.4 times.

It is worth noting that the speedup of computations in the time depends on the time of computing the values of the criteria of the MCO problem being solved. This time is relatively small in the test optimization problems, but it can be essential when solving the applied problems in various scientific and technical applications. As a result, along with the evaluation of the achieved speedup of computations in the time, it is reasonable to evaluate the speedup of computations with respect to the reduction of the number of iterations performed in the course of computations. The results of experiments performed to evaluate such speedup are presented in Table \ref{tab:03}.

\begin{table}[htbp]
\centering
\caption{Comparison of the number of iterations when solving the six-dimensional bicriterial MCO problems}
\label{tab:03}
\begin{tabular}{ccccccc}
\hline
\textbf{Nodes} & \textbf{P} & \textbf{Th} & \textbf{P*Th} & \textbf{Core type} & \textbf{Iterations} & \textbf{Speedup} \\ \hline
1              & 1          & 16          & 16            & CPU                                                           & 12 279 179.8        & 1.0              \\
16             & 16         & 16          & 256           & CPU                                                           & 808 858.8           & 15.2             \\
16             & 32         & 4 032       & 129 024       & GPU                                                           & 3 086.5             & 3 978.4          \\
16             & 64         & 2 016       & 129 024       & GPU                                                           & 2 426.9             & 5 059.6          \\
16             & 128        & 1 008       & 129 024       & GPU                                                           & 2 910.2             & 4 219.4          \\
32             & 128        & 2 016       & 258 048       & GPU                                                           & 1 581.5             & 7 764.4          \\
32             & 256        & 1 008       & 258 048       & GPU                                                           & 2307.5              & 5 321.4          \\ \hline
\end{tabular}
\end{table}

The results of experiments presented above demonstrate the speedup of computations with respect to the reduction of the number of performed global search iterations to be considerable. Thus, when employing 16 computer nodes, the speedup can be more than 5000 times whereas when employing 32 computer nodes -- more than 7700 times.

\section{Conclusion}
In the present paper, an efficient approach for solving the complex multicriterial optimization problems, in which the criteria of optimality can be the multiextremal ones, and the computing of the criteria values can require a large amount of computations has been proposed. The proposed approach is based on the reduction of the multicriterial problems to the scalar optimization ones by means of the minimax convolution of the partial criteria, the dimensionality reduction with the use of Peano space-filling curves, and the application of the efficient information-statistical methods of global optimization.

The opportunity of the large-scale parallel computations is provided by application of the block multistep dimensionality reduction scheme and by the use of the GPUs with many thousands computational cores. The results of numerical experiments have shown the developed approach to allow reducing the computational costs of solving the multicriterial optimization problems considerably -- by hundreds and thousands times.

The results of the performed experiments have demonstrated the developed approach to be a promising one and to require further investigations. First of all, it is necessary to continue carrying out the numerical experiments on solving the multicriterial optimization problems at larger number of the partial criteria of efficiency and for larger dimensionality of the optimization problems being solved.

\section*{Acknowledgements}
This research was supported by the Russian Science Foundation, project No 16-11-10150 ''Novel efficient methods and software tools for time-consuming decision-making problems using supercomputers of superior performance''.

\begin{thebibliography}{8}

\bibitem{c2} Marler, R.T., Arora, J.S.: Multi-Objective Optimization: Concepts and Methods for Engineering,  VDM Verlag  (2009)
\bibitem{c3} Ehrgott, M.: Multicriteria Optimization,  Springer, 2005. (2nd ed., 2010)
\bibitem{c4} Collette, Y., Siarry, P.: Multiobjective Optimization: Principles and Case Studies (Decision Engineering),  Springer (2011)
\bibitem{c5} Pardalos, P.M., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-Convex Multi-Objective Optimization,  Springer (2017)
\bibitem{c7} Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications,  Surv. Math. Ind. \textbf{11}, pp. 1--42 (2005)
\bibitem{c9} Cho, J.-H., Wang, Ya., Chen, I.-R., Chan, K.S., Swami, A.: A Survey on Modeling and Optimizing Multi-Objective Systems,  IEEE Communications Surveys {\&} Tutorials \textbf{19}(3), pp. 1867--1901 (2017) 
\bibitem{c10}	Eichfelder, G.: Scalarizations for adaptively solving multi-objective optimization problems,  Comput. Optim. Appl. \textbf{44}, pp. 249--273 (2009) 
\bibitem{c11}	Strongin, R., Sergeyev, Ya.: Global optimization with non-convex constraints. Sequential and parallel algorithms,  Kluwer Academic Publishers, Dordrecht, 2000 (2nd ed. 2013, 3rd ed. 2014).
\bibitem{c12}	Strongin, R., Gergel, V., Grishagin, V., Barkalov, K.: Parallel computations for global optimization problems,  Moscow State University Press (2013) (in Russian)
\bibitem{c13}	Sergeyev, Y.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves,  Springer (2013)
\bibitem{c14}	Gergel V.P., Kozinov, E.A.: Accelerating Parallel Multicriterial Optimization Methods Based on Intensive Using of Search Information,  Procedia Computer Science, \textbf{108}, pp. 1463-1472 (2017)
\bibitem{c15}	Gergel, V., Kozinov, E.: Parallel computing for time-consuming multicriterial optimization problems,  Lecture Notes in Computer Science, \textbf{10421}, pp. 446--458 (2017).  DOI: 10.1007/978-3-319-62932-2\_43
\bibitem{c16}	Gergel, V., Kozinov, E.: Efficient methods of multicriterial optimization based on the intensive use of search information,  Springer Proceedings in Mathematics and Statistics, \textbf{197}, pp. 27--45 (2017). DOI: 10.1007/978-3-319-56829-4\_3
\bibitem{c17}	Gergel, V., Kozinov, E.: An approach for parallel solving the multicriterial optimization problems with non-convex constraints,  Communications in Computer and Information Science, \textbf{793}, pp. 121-135 (2017). DOI: 10.1007/978-3-319-71255-0\_10
\bibitem{c18}	Cai, Y., See, S.: GPU Computing and Applications,  Springer (2015) 
\bibitem{c19}	Ferreiro, A.M., Garcia, J.A., Lopez-Salas, J.G., Vazquez, C.: An efficient implementation of parallel simulated annealing algorithm in GPUs,  J. Glob. Optim., \textbf{57}(3), pp. 863--890 (2013)
\bibitem{c20}	Zhu, W.: Massively parallel differential evolution--pattern search optimization with graphics hardware acceleration: an investigation on bound constrained optimization problems,  J. Glob. Optim. \textbf{50}(3), pp. 417--437 (2011)
\bibitem{c21}	Garcia-Martinez, J.M., Garzon, E.M., Ortigosa, P.M.: A GPU implementation of a hybrid evolutionary algorithm: GPuEGO,  J. Supercomput (2014). DOI:10.1007/s11227-014-1136-7
\bibitem{c22}	Langdon, W.B.: Graphics processing units and genetic programming: an overview,  Soft. Comput. \textbf{15}(8), pp. 1657--1669 (2011)
\bibitem{c27}	Locatelli, M., Schoen, F.: Global Optimization: Theory, Algorithms, and Applications,  SIAM (2013)
\bibitem{c28}	Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization,  Princeton University Press (2016)
\bibitem{c29}	Gergel, V.P., Strongin, R.G.: Parallel Computing for Globally Optimal Decision Making,  Lecture Notes in Computer Science, \textbf{2763}, pp. 76--88 (2003)
\bibitem{c30}	Gergel, V.P., Kuzmin, M.I., Solovyov, N.A., Grishagin, V.A.: Recognition of surface defects of cold-rolling sheets based on method of localities,  International Review of Automatic Control \textbf{8}(1), pp. 51--55 (2015)
\bibitem{c32}	Modorskii, V.Y., Gaynutdinova, D.F., Gergel, V.P., Barkalov, K.A.: Optimization in design of scientific products for purposes of cavitation problems,  AIP Conference Proceedings, \textbf{1738}, pp. 400013 (2016). DOI: 10.1063/1.4952201
\bibitem{c33} Grishagin, V., Israfilov, R., Sergeyev, Y.: Convergence conditions and numerical comparison of global optimization methods based on dimensionality reduction schemes,  Applied Mathematics and Computation, \textbf{318}, pp. 270--280 (2018). DOI: 10.1016/j.amc.2017.06.036
\bibitem{c34}	Sergeyev, Y., Grishagin, V.: Parallel asynchronous global search and the nested optimization scheme,  J. Comput. Anal. Appl., \textbf{3}(2), pp. 123--145 (2001)
\bibitem{c36}	Barkalov, K.A., Gergel, V.P.: Multilevel scheme of dimensionality reduction for parallel global search algorithms,  Proceedings of the 1st International Conference on Engineering and Applied Sciences Optimization, pp. 2111--2124 (2014)
\bibitem{c38} Gergel, V., Lebedev, I.: Heterogeneous Parallel Computations for Solving Global Optimization Problems,  Procedia Computer Science, \textbf{66}, pp. 53--62 (2015)
\bibitem{c39}	Gergel, V.P., Sidorov, S.V.: A Two-Level Parallel Global Search Algorithm for Solving Computationally Intensive Multi-extremal Optimization Problems,  Lecture Notes in Computer Science, \textbf{9251}, pp. 505--515 (2015)
\bibitem{c40}	Gergel, V.: An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems,  2nd International Conference on Mathematics and Computers in Sciences and in Industry, MCSI, \textbf{7423935}, pp. 13--18 (2016)
\bibitem{c41}	Evtushenko, Yu.G., Posypkin, M.A.: A deterministic algorithm for global multi-objective optimization,  Optimization Methods \& Software, \textbf{29}(5), pp. 1005--1019 (2014)
\bibitem{c42}	{\v Z}ilinskas, A., {\v Z}ilinskas, J.: Adaptation of a one-step worst-case optimal univariate algorithm of bi-objective Lipschitz optimization to multidimensional problems,  Commun Nonlinear Sci Numer Simulat, \textbf{21}, pp. 89--98 (2015)
\bibitem{c43}	Gaviano, M., Kvasov, D.E., Lera, D., Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization,  ACM Transactions on Mathematical Software \textbf{29}(4), pp. 469--480 (2003)

\end{thebibliography}

\end{document}

### 2018-12-11 2019/ICCS_MCO/ICCS_2019.tex

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%%!PS-Adobe-3.0 EPSF-3.0
%%%BoundingBox: 19 19 221 221
%%%CreationDate: Mon Sep 29 1997
%%%Creator: programmed by hand (JK)
%%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[runningheads]{llncs}
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
%
\usepackage{graphicx}
\usepackage{mathtools,amssymb}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{A Highly Parallel Approach for Solving Computationally Expensive Multicriteria Optimization Problems
\thanks{This work has been supported by Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using superior-performance supercomputers.''}
}

\titlerunning{A Highly Parallel Approach for Solving MCO Problem}        % if too long for running head

\author{Victor Gergel\orcidID{0000-0002-4013-2329} \and Evgeny Kozinov \orcidID{0000-0001-6776-0096}}

\authorrunning{V. Gergel et al.} % if too long for running head

\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
              \email{gergel@unn.ru, evgeny.kozinov@itmm.unn.ru}
}

\maketitle

\begin{abstract}
In this paper, a highly parallel approach for solving multicriteria optimization problems is proposed. The considered approach is based on the reduction of the multicriterial problems to the global optimization ones using the minimax convolution of the partial criteria, the dimensionality reduction with the use of the Peano space-filling curves, and the application of the efficient parallel information-statistical global optimization methods. The required computations can be time-consuming since functions representing individual criteria can be multiextremal and computationally expensive. The proposed approach comprises two different schemes for efficient parallel computations on high performance systems with shared and distributed memory and with a large number of computational units. The computational efficiency is achieved by storing all the computed criteria values and their intensive reuse for finding new solutions. The results of numerical experiments have demonstrated that this approach allows to reduce the computational costs of solving multicriteria optimization problems by a factor between 10 and 100. 
\keywords{Multicriteria optimization \and Parallel computing \and Dimensionality reduction \and Criteria convolution \and Global optimization algorithm \and Computational complexity}
\end{abstract}

\section{Introduction}
\label{sec:1}

The multicriteria optimization (MCO) problems arise in a general formulation of a decision making problem. Reviews of the state of the art in multicriteria optimization can be found in \cite{c1,c2,c3,c12} and reviews of relevant techniques and applications in \cite{c4,c5,c6,c7,c13,c50}.

At the same time, the MCO problems are some of the most complex ones. To solve an MCO problem compromise decisions need to be found and having to find a representative set of those increases the computational complexity of solving the MCO problems. Bearing in mind that finding even a single efficient solution may require a large amount of computations, obtaining the whole Pareto set or even a limited subset of efficient decisions becomes a problem of high computational complexity. In order to solve such problems huge computational capabilities of high-performance systems are required.

The structure of this paper is as follows. Section \ref{sec:2} presents a general statement of a multicriteria optimization problem. A scheme of parallel computations to simultaneously solve a set of multicriteria global optimization problems is proposed in Section \ref{sec:3}. A parallel algorithm of multicriteria global search is presented in Section \ref{sec:4}. To demonstrate the proposed techniques, Section \ref{sec:5} presents results of numerical experiments.

\section{Statement of a Multicriteria Optimization Problem}
\label{sec:2}
The problem of multicriteria (or multi-objective) optimization (MCO) can be defined as follows:
\begin{equation}
f(y) =(f_1 (y),f_2 (y),\dots,f_s (y) )\to min,y\in D,	
\label{eq:1}
\end{equation}
where $y = (y_1,y_2,\dots,y_N)$ is a vector of varied parameters describing a system being designed, $N$ is the dimensionality of the optimization problem, and $D$ is the search domain being an $N$-dimensional hyperinterval
\begin{equation*}
D= \{ y \in R^N: a_i \leq y_i \leq b_i,1 \leq i \leq N \}
\end{equation*}
with given boundary vectors $a$ and $b$. Without a loss of generality, the efficiency criteria values in the problem (\ref{eq:1}) are assumed to be non-negative, and the decrease of these ones corresponds to the increase of the efficiency of the considered solutions $y \in D$. In addition it is assumed the efficiency criteria $f_i(y)$, $1 \leq i \leq s$ can be multiextremal, and the obtaining of the criteria values at the points of the search domain $y \in D$ can require a large amount of computations. It is assumed that the efficiency criteria $f_i(y)$, $1 \leq i \leq s$ satisfy the Lipschitz condition
\begin{equation}
|f_i (y') - f_i (y'') | \leq L_i ||y' - y''||, y', y'' \in D, 1\leq i\leq s	
\label{eq:2}
\end{equation}
where $L_i$, $1 \leq i \leq s$ are the Lipschitz constants for the corresponding efficiency criteria $f_i(y)$, $1 \leq i \leq s$, and $||.||$  denotes the norm in the $R^N$ space.

As \textit{a solution} of the MCO problem, any efficient decision  is considered. In a general case, it is required to find the whole set of Pareto-optimal solutions $PD(f,D)$, i.e. \textit{the complete solution of an MCO problem}.

A general approach to solving the MCO problem used in this paper consists in the replacement of solving the original MCO problems by solving a series of simpler optimization problems: 
\begin{equation}\label{eq:3}
\begin{split}
\min{(\phi(x) = F(\lambda, y(x)), y(x) \in D, x \in [0,1])},\\
F(\lambda,y(x))=\max⁡{(\lambda_i f_i(y(x)), 1 \leq i \leq s)},\\
\lambda=(\lambda_1,\lambda_2,\dots,\lambda_s) \in \Lambda \subset R^s : \sum_{i=1}^s{\lambda_i}=1, \lambda_i \geq 0, 1 \leq i \leq s
\end{split}
\end{equation}
where $F(\lambda,y(x))$ is the minimax convolution of the efficiency criteria of the MCO problem \cite{c2,c6,c19} with the use of the vector of the convolution coefficients $\lambda \in \Lambda$ and $y(x)$ is a continuous and unambiguous mapping of the interval [0,1] onto an $N$-dimensional search domain $D$ -- see, for example, \cite{c7,c8}.

The necessity and sufficiency of this approach for solving MCO problems is a key property of the minimax convolution. The result of the minimization of $F(\lambda,y)$ leads to obtaining an efficient variant\footnote{More precisely, the minimization of $F(\lambda,y)$ can lead to the obtaining of the weakly-efficient variants (the set of the weakly-efficient decisions includes the Pareto domain).} of the MCO problem and vice versa, any efficient variant of an MCO problem can be obtained as a result of the minimization of $F(\lambda,y)$ at corresponding values of the convolution coefficients $\lambda_i$, $1 \leq i \leq s$. In order to obtain several efficient decisions (or in order to evaluate the whole Pareto domain) the problem (\ref{eq:3}) should be solved for the corresponding set of values of the vector $\lambda \in \Lambda$.

The dimensionality reduction applied in (\ref{eq:3}) allows to replace solving multidimensional MCO problems by the optimization of one-dimensional functions $F(\lambda,y(x))$ which satisfy the uniform H\"{o}lder condition i.e.
\begin{eqnarray} \label{eq:04}
|F(\lambda, y(x')) - F(\lambda, y(x''))| \leq H |x' - x''|^{1/N}, x', x'' \in [0,1], 
\end{eqnarray}
where the constant $H$ is defined as $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant of the function $F(\lambda,y)$ and $N$ is the dimensionality of the optimization problem (\ref{eq:1}).

\section{Parallel Computations for Solving the Multicriteria Global Optimization Problems}
\label{sec:3}
The scalarization of the vector criterion allows reducing the solving of the MCO problem (\ref{eq:1}) to solving a series of the multiextremal problems (\ref{eq:3}). Therefore, the problem of development of the methods for solving the MCO problems is resolved by the possibility of an extensive use of efficient parallel global optimization algorithms.

Multiextremal optimization is a research area being actively developed -- its state of the art is presented, for example, in \cite{c7,c9,c10}. The information-statistical theory of global optimization is considered to be one of the promising approaches \cite{c7}. HPC systems are used widely for solving time-consuming global search problems \cite{c7,c11,c14,c15}.

The proposed approach to the use of parallel computations for solving time-consuming global optimization problems is based on the following main statements:
\begin{itemize}

\item The parallelism of the performed computations is provided by means of simultaneous computing the values of the efficiency criteria $f_i (y)$, $1 \leq i \leq s$ at several points of the search domain $D$. Such an approach provides the parallelization of the most computation-costly part of global optimization and is a general one -- it can be applied with many global optimization methods for various global optimization problems.

\item The parallel computations are provided by means of simultaneous solving of several global optimization problems (\ref{eq:3}) for varying values of the coefficients $\lambda_i$, $1 \leq i \leq s$. For solving the problems of the family (\ref{eq:3}), a set of computational nodes of the high performance systems with distributed memory can be applied.

\item The optimization data obtained in the course of parallel computations is exchanged between all employed processors because of the information compatibility of the global optimization problems of the family (\ref{eq:3}).
\end{itemize}

All these statements will be considered in more details below.

\subsection{General Scheme of Parallel Computations }
\label{ssec:31}

As mentioned above, when solving a multicriteria optimization problem (\ref{eq:1}), solving a series of scalar problems (\ref{eq:3}) with varying coefficients of the minimax convolution of the efficiency criteria may be required in order to find several different efficient decisions:
\begin{equation}
\Phi(y)=\{\phi_1 (y),\dots,\phi_q(y) \}, \phi_i(y)=F(\lambda_i,y), 1 \leq l \leq q.
\label{eq:5}
\end{equation}

Such problems can be solved sequentially by various global optimization methods. Alternatively, these problems can be solved simultaneously using several processors. It is important to note that the obtained family of the one-dimensional problems $\Phi(y)$ is an information-dependent one because the values of the optimized functions computed for any problem $\phi_l (y)$, $1 \leq l \leq q$ can be transformed to the values of all the remaining problems of the family without time-consuming recalculations of the efficiency criteria values $f_i (y)$, $1 \leq i \leq s$ (see Section \ref{ssec:32}).

The information compatibility of the problems from the family (\ref{eq:5}) allows to propose the following unified scheme of parallel computations:
\begin{enumerate}
	\item The family of the information-linked problems $\Phi(y)$ in (\ref{eq:5}) is distributed among the processors of the computing system. 
  \item For solving the optimization problems (\ref{eq:5}), a global optimization method implemented on each processor should be updated following the following rules:
  \begin{enumerate}
		\item Upon completing the iteration for any problem $\phi_l(y)$, $1 \leq l \leq q$ at any point $y' \in D$, the point $y'$ with the particular criteria values $f_i(y)$, $1 \leq i \leq s$ computed at this point should be transferred to all employed processors.
	  \item Prior to the start of the next global search iteration, the method should check the queue of the received messages; if there are any data in the queue, the received information should be included into the search information.
	\end{enumerate}
\end{enumerate}

The possibility of the asynchronous data transfer is a key feature of such a scheme of the parallel computations. It is important to note that this scheme does not depend on any of the single control nodes, and the number of computational nodes can vary in the course of global optimization.

\subsection{Accumulation and Reuse of Calculated Optimization Data}
\label{ssec:32}

For solving the stated optimization problems, the values of the efficiency criteria $f^i=f(y^i)$ at the points $y^i$, $1 \leq i \leq k$ of the search domain $D$ are computed. The search information obtained as a result of these computations can be represented in the form of the \textit{Search Information Set} (SIS):
\begin{equation}\label{eq:6}
\Omega_k=\{(y^i,f^i=f(y^i ) )^T:1 \leq i \leq k\}.
\end{equation}
As a result of the dimensionality reduction, the search information $\Omega_k$ from (\ref{eq:6}) can be transformed into the \textit{Matrix of the Search State} (MSS) 
\begin{equation}\label{eq:7}
A_k=\{(x_i,z_i,l_i )^T:1 \leq i \leq k\}
\end{equation}
where $x_i$, $1 \leq i \leq k$ are the reduced points of the executed global search  iterations\footnote{The lower indices denote the increasing order of the coordinate values of the points $x_i$, $1 \leq i \leq k$.}, $z_i$, $1 \leq i \leq k$ are the values of the scalar criterion of the current optimization problem (\ref{eq:3}), $l_i$, $1 \leq i \leq k$ are the indices of the optimization iterations where the points $x_i$, $1 \leq i \leq k$ were computed.

The matrix of the search state can be used by the optimization algorithms in order to improve the efficiency of the global search -- selecting the points for the scheduled iterations can be performed taking into account the results of all previously executed computations.

\subsection{Parallel Computations on Multiprocessors with Shared Memory }
\label{ssec:33}

Within the proposed approach, for solving the reduced one-dimensional multiextremal optimization subproblems (\ref{eq:3}) it is proposed to use the well-known Parallel Algorithm of Global Search (PAGS) developed within the framework of the information-statistical theory of the multiextremal optimization \cite{c7}. This method has a good theoretical background and has demonstrated its efficiency as compared to other global search algorithms (see also the results of numerical experiments in Section \ref{sec:4}).

For completeness, let us briefly discuss the general computational scheme of PAGS, which consists in the following.

Let $p$ be the number of employed parallel computers (processors or cores) of a computational system with shared memory. The initial two iterations of the algorithm are performed at the boundaries of the interval $x^0=0$, $x^1=1$. Apart from these boundary points, the algorithm should perform additional iterations at the points $x^i$, $1<i \leq p$ which can be defined \textit{a priori} or computed by any auxiliary computational procedure. Then, let $(k>p)$ global search iterations be completed, at each of which the computation of the value of the minimized function $\phi(x)$ from (\ref{eq:3}) (hereafter called \textit{a trial}) has been performed. The choice of the points for trials performed within the next iteration in parallel is determined by the following rules.

\textit{Rule 1.} Renumber the trial points of the completed search iterations with the lower indices in the order of increasing coordinate values
\begin{eqnarray} \label{eq:08}
	0=x_0<x_1<\dots<x_i<\dots<x_k=1.
\end{eqnarray}

\textit{Rule 2.} Compute the current estimate of the H\"{o}lder constant of the function $\phi(x)$:
\begin{eqnarray} \label{eq:09}
	m = 
 \begin{cases}
   r M, & M > 0 \\
   1,   & M = 0
 \end{cases} ,
 M = \max_{1 \leq i \leq k} {\frac{|z_i - z_{i-1}|} {\varrho_i} }
\end{eqnarray}
where $z_i=\phi(x_i)$, $\varrho_i=\sqrt[N]{x_i-x_{i-1}}$, $1 \leq i \leq k$. The constant $r$, $r>1$, is the \textit{reliability parameter} of the algorithm.

\textit{Rule 3.} Compute \textit{the characteristic} $R(i)$ for each interval $(x_{i-1},x_i)$, $1 \leq i \leq k$ according to the expression 
\begin{eqnarray} \label{eq:10}
R(i) = \varrho_i + \frac{(z_i - z_{i-1})^2}{m^2 \varrho_i} - 2 \frac{(z_i + z_{i-1})}{m}, 1 \leq i \leq k.
\end{eqnarray}

\textit{Rule 4.} Arrange the characteristics of the intervals $(x_{i-1},x_i)$, $1 \leq i \leq k$ obtained according to (\ref{eq:10}) in the decreasing order 
\begin{eqnarray} \label{eq:11}
R(t_1) \geq R(t_2)\geq \dots \geq R(t_{k-1}) \geq R(t_k) 
\end{eqnarray}
and select $p$ intervals with the indices $t_j$, $1 \leq j \leq p$ having the maximum values of the characteristics.

Rule 5. Perform new trials at the points $x^{k+j}$, $1 \leq j \leq p$ placed in the intervals with the maximum characteristics from (\ref{eq:11}) according to the expressions

\begin{eqnarray} \label{eq:12}
	x^{k+j} = \frac{x_{t_j} + x_{t_j-1}}{2}
	- sign(z_{t_j} - z_{t_j-1}) \frac{[\frac{|z_{t_j} - z_{t_j-1}|}{m}]^N}{2r} , 1\leq j \leq p.
\end{eqnarray}

\textit{Termination condition} for the algorithm, according to which the execution of the algorithm is terminated, consists of checking the lengths of the intervals in which the scheduled trials are performed with respect to a required \textit{accuracy} of the problem solution, i. e.

\begin{eqnarray} \label{eq:13}
\varrho_{t_j} \leq \varepsilon, 1 \leq t_j \leq p.
\end{eqnarray}

As the current estimate of the optimization problem solution, the minimum computed value of the optimization function is accepted i. e.
\begin{eqnarray} \label{eq:14}
z_k^*=\min⁡{\{z_i:1\leq i \leq k\}}.
\end{eqnarray}

Within the framework of the proposed approach, PAGS is applied to solving the MCO problem (\ref{eq:1}) according to the scheme (\ref{eq:3}). The availability of SIS from (\ref{eq:6}) allows transforming the results of the previous computations to the values of a next optimization problem (\ref{eq:3}) without any time-consuming computations of the efficiency criteria values for new values of the convolution coefficients, i.e.
\begin{eqnarray} \label{eq:15}
z_i'=\max{⁡(\lambda_j' f_i^j,1\leq j \leq s)},1 \leq i \leq k.
\end{eqnarray}

As a result, all the search information accumulated so far can be employed for continuing the computations. In general, the reuse of the search information will provide a smaller and smaller amount of computations for solving every successive optimization problem (see the results of numerical experiments in Section \ref{sec:4}).

The method obtained as a result of such extension is called hereafter Parallel Multicriteria Global Algorithm (PMGA) for multiprocessors with shared memory. The PMGA algorithm in combination with a general scheme of the parallel computations presented in Section \ref{ssec:31} will be referred to as the Multilevel Parallel Multicriteria Global Algorithm (MPMGA) for high-performance computational systems with shared and distributed memory.

Theoretical properties of MPMGA can be established by a theoretical analysis of the sequential one-dimensional variant of the method, i.e. the MGA algorithm. Here is a summary (without a proof) of the theoretical results from \cite{c16}.

\textbf{Theorem 1.} When switching from solving the optimization subproblem (\ref{eq:3}) with the convolution coefficients $\lambda' \in \Lambda$ to solving the next subproblem (\ref{eq:3}) with the coefficients $\lambda'' \in \Lambda$, the value of the function $F(\lambda'',y(x))$ at the point of the global minimum differs from the estimated minimum value $z_k^*$ from (\ref{eq:14}) obtained by minimizing the function $F(\lambda',y(x))$, by no more than the bounded value 
\begin{eqnarray} \label{eq:16}
|z_k^*-F(\lambda'',y(x^*))| \leq \frac{\alpha H \varepsilon}{2} + \delta, \alpha>1,
\end{eqnarray} 
where $\delta$ is the maximum possible change of the function $\phi(x)$ (being minimized) from one subproblem (\ref{eq:3}) to the next:

\begin{equation} \label{eq:17}
\begin{split}
\delta=\Delta \phi_{max}=\max{\{\Delta \lambda_{max} f_i^{max}, 1 \leq i \leq s \}},\\
\Delta \lambda_{max}=\max{\{|\lambda_i'-\lambda_i'' |, 1 \leq i \leq s\}},	\\
f_i^{max} = \max {\{f_i (y), y \in D\} }, 1 \leq i \leq s.
\end{split}
\end{equation} 
and for the value $m$ from (\ref{eq:09}) the following inequality is satisfied
\begin{equation*}
m \geq H\left(1+\sqrt{\frac{\alpha+4\beta(1+\beta)}{\alpha-1}} \right), \beta = \frac{2r\delta}{H\varepsilon(r-1)}
\end{equation*} 
with $z_k^*$ from (\ref{eq:14}), $\phi(x^*)$ from (\ref{eq:3}), $\varepsilon$ from (\ref{eq:13}), $r$ from (\ref{eq:09}) and $H$ from (\ref{eq:04}).

This statement means that if the deviation $\Delta \phi_{max}$ from (\ref{eq:17}) of the estimates of the minimum values of $F(\lambda'',y(x))$ is acceptable, then solving the subproblem (\ref{eq:3}) with the new convolution coefficients $\lambda'' \in \Lambda$ does not require any additional optimization iterations because the estimates of the minimum value of the function $F(\lambda'',x)$ can be obtained according to (\ref{eq:16}) using the values $z_i$, $1 \leq i \leq k$, located within the search information $A_k$ from (\ref{eq:7}).

As a result, two different schemes for selecting various convolution coefficients $\lambda \in \Lambda$ can be proposed, with whose use the required conditions of the Theorem 1 will be satisfied. In the first scheme, initially a sparse grid can be built on the set $\Lambda$, so that when setting new values of $\lambda$, the values that are close to the ones already established on the grid built in $\Lambda$ earlier could always be found. In the second scheme, new values of $\lambda$ could be chosen by small perturbations of the values used earlier (this method is often used in solving practical MCO problems when the decision maker wants to refine the estimates of the efficient solutions obtained earlier). It is worth noting that the reuse of the search information could be beneficial even with significant differences in the chosen values of the coefficients $\lambda$ from the previously established ones, see the results of numerical experiments in Section \ref{sec:4}.

\section{Results of Numerical Experiments}
\label{sec:4}

The numerical experiments have been carried out on the \textit{Lobachevsky} supercomputer at State University of Nizhni Novgorod. Each supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz octa-core processors, 64 Gb RAM.

To evaluate the efficiency the proposed approach, the MGA\footnote{MGA is the sequential implementation of the parallel MPMGA method} algorithm was compared with a number of other popular multicriteria optimization methods. In \cite{c20} MGA was tested against four multicriteria optimization methods: the Monte-Carlo (MC) method where the trial points are selected within the search domain $D$ randomly and uniformly, the genetic algorithm SEMO from the PISA library \cite{c21}, the Non-Uniform Coverage (NUC) method \cite{c17} and the Bi-objective Lipschitz Optimization (BLO) method \cite{c18}.

In \cite{c16} a series of numerical experiments were executed to compare MGA with four other MCO methods considered in \cite{c22}, namely the Linear Combination Method (LCM), the Multi-Objective Genetic Algorithm (MOGA), the Global Criterion Method (GCM), the $\varepsilon$-Constraint Method (ECM).

The results of the numerical experiments have demonstrated that MGA has a considerable advantage compared to the considered multicriteria optimization methods even when solving the relatively simple MCO problems.

In this paper, we present the results of computational experiments to evaluate the efficiency of the MPMGA algorithm. The computations have been carried out for 100 of four-dimensional MKO problems with ten criteria, i.e. $N = 4$, $s = 10$. The criteria in these problems were obtained by the GKLS generator \cite{c23} that generates multiextremal optimization problems with a priori known properties: the number of local minima, the size of their attraction domains, the point of the global minimum, etc.

In these experiments 100 multicriteria problems has been solved. To obtain the Pareto domain approximation, each problem has been solved for 50 coefficients $\lambda$ uniformly distributed in $\Lambda$ from (\ref{eq:3}). The obtained results were averaged over the number of solved MCO problems. The accuracy $\varepsilon$ from (\ref{eq:13}) was 0.025 and the reliability parameter $r$ from (\ref{eq:09}) was 5.6.

The following approach was used to evaluate the accuracy of the Pareto domain approximation calculated for each of the 100 solved MCO problems. First of all, for the domain $D$ from (\ref{eq:1}) a uniform grid $D_\delta$ was constructed with the step $\delta = 0.01$ for each coordinate. On this grid $D_\delta$, the minimum values were found for all 50 subproblems $F(\lambda,y(x))$  from (\ref{eq:3}) generated in accordance with the chosen set of coefficients $\lambda \in \Lambda$ -- the resulting set $y\in PD_\delta (f,D)\subset D_\delta$ was taken as the $\delta$-approximation of the Pareto domain of the MCO problem being solved.

To solve each subproblems $F(\lambda,y(x))$ in (\ref{eq:3}), the termination condition (\ref{eq:13}) was replaced by the termination condition by the condition
\begin{equation*}
F(\lambda,y(x^{k+j})) \leq F(\lambda,y^* )+\varepsilon,1 \leq j \leq p,  F(\lambda,y^*) = \min_{y\in D_\delta}⁡{F(\lambda,y)}.
\end{equation*}

After solving subproblems (\ref{eq:3}) for 50 values of $\lambda$ a numerical approximation of the Pareto domain $PD_{\Omega_k}(f,D)$ is constructed using the search information accumulated in $\Omega_k$ from (\ref{eq:6}) as result of the solution of the MCO problem. The resulting estimation of the accuracy of the calculated approximation $PD_{\Omega_k}(f,D)$ is defined as the closeness to the $\delta$-approximation $PD_\delta(f,D)$ of the Pareto domain of the MCO problem 
\begin{equation}\label{eq:18}
\Delta=\frac{1}{M} \sum_{y\in PD_\delta(f,D)}{d(y)}, d(y)=\min_{\overline{y}\in PD_{\Omega_k}(f,D)}{⁡\|y-\overline{y}\|},
\end{equation}
where $\|.\|$ denotes the norm in the $R^N$ space, and $M$ is the number of points used in the approximation $PD_{\delta}(f,D)$.

Results of the numerical experiments are presented in Table \ref{tab:01}. The first two columns in Table \ref{tab:01} show the numbers of the processors (\textit{P}) and of the parallel computational cores on each processor (\textit{Q}) employed. The third column (\textit{P*Q}) contains the total number of cores employed. In the fourth (\textit{Iters}) column the average numbers of iterations spent on solving the problems for the corresponding numbers of the different coefficients $\lambda$ from (\ref{eq:3}) are given. In the fifth column (\textit{Nums}) the average numbers of the calculated solutions used to establish the approximation $PD_{\Omega_k}(f,D)$ are shown. In the sixth column (\textit{Approx}) the accuracy $\Delta$ of the $\delta$-approximation $PD_\delta(f,D)$ is presented.

The last two columns show the speedup \footnote{Due to the initial assumption that the computational complexity of multicriteria optimization problems is determined by the complexity of calculations of the criteria values, the speedup is defined as the reduction of the number of  executed iterations.} of the parallel computations obtained with the use of the search information (\textit{S}1) and without (\textit{S}2).

\begin{table}[ht]
\centering
\caption{Numerical results for solving four-dimensional problems with ten criteria}
\label{tab:01}
\begin{tabular}{cccccccc}
\hline
P     & Q     & P*Q    & Iters       & Nums        & Approx   & S1         & S2       \\ \hline
\multicolumn{8}{c}{\textbf{Computations without the reuse of the search information}} \\ \hline
1     & 1     & 1      & 82,244.11   & 26,014.27   & 0.05     &            & 1,0      \\ \hline
\multicolumn{8}{c}{\textbf{Computations with the reuse of the search information}}    \\ \hline
1     & 1     & 1      & 23,791.81   & 8,185.07    & 0.08     & 1.00       & 3.5      \\
5     & 1     & 5      & 4,977.63    & 8,077.38    & 0.08     & 4.78       & 16.5     \\
1     & 20    & 20     & 1,223.45    & 8,452.35    & 0.08     & 19.45      & 67.2     \\
25    & 1     & 25     & 1,063.22    & 8,704.86    & 0.08     & 22.38      & 77.4     \\
1     & 40    & 40     & 616.37      & 8,416.06    & 0.07     & 38.60      & 133.4    \\
50    & 1     & 50     & 594.46      & 9,561.09    & 0.07     & 40.02      & 138.4    \\
5     & 20    & 100    & 252.65      & 8,510.26    & 0.07     & 94.17      & 325.5    \\
5     & 40    & 200    & 131.68      & 8,914.21    & 0.07     & 180.68     & 624.6    \\
25    & 20    & 500    & 66.54       & 11,022.32   & 0.07     & 357.56     & 1236.0   \\
25    & 40    & 1000   & 36.81       & 12,226.82   & 0.07     & 646.34     & 2234.3   \\
50    & 20    & 1000   & 37.94       & 12,552.88   & 0.07     & 627.09     & 2167.7   \\
50    & 40    & 2000   & 20.98       & 14,399.94   & 0.07     & 1,134.02   & 3920.1   \\ \hline
\end{tabular}
\end{table}

The obtained results of experiments demonstrate that even simple reuse of the search information allows reducing the total amount of computations by the factor of 3.5 without the use of additional computational resources. When 1000 computational cores were used, the obtained speedup varied from 627 to 646. If 2000 computational cores are used, the speedup with the reuse of the search information reaches 1134. The overall speedup in this case relative to the initial algorithm without the reuse of the search information was 3920.

\section{Conclusion}
\label{sec:5}
In this paper, a new computationally efficient approach is proposed that allows a parallel solution of time-consuming multicriteria optimization problems where individual criteria can be multiextremal and their evaluation may require a large computing effort. The main feature of this approach is its efficient handling of the computational complexity of solving such multicriteria optimization problems. Improvements of the efficiency resulting in a significant reduction of the required computations have been achieved by an intensive use of the search information obtained in the course of computations. Within the framework of this approach, methods for a re-use of the already available search information for a currently handled scalar nonlinear pro-gramming problem have been proposed. Such search information is then used by the optimization methods for the adaptive planning of the iterative process of global search.

Results of numerical experiments demonstrate that such an approach allows reducing the computation costs of solving multicriteria optimization problems by a factor between tens and hundreds.

In conclusion, it was shown that the developed approach is a promising one that needs a further investigation. It is important to continue numerical experiments of solving multicriteria optimization problems with a larger number of criteria and of a larger dimensionality.

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

% Non-BibTeX users please use
\begin{thebibliography}{}
%
% and use \bibitem to create references. Consult the Instructions
% for authors for reference list style.
%
\bibitem{c1} 	Miettinen, K.: Nonlinear Multiobjective Optimization. In: Springer (1999)
\bibitem{c2} 	Ehrgott, M.: Multicriteria Optimization. In: Springer (2nd ed., 2010)
\bibitem{c3} 	Collette, Y., Siarry, P.: Multiobjective Optimization: Principles and Case Studies (Decision Engineering). In: Springer (2011)
\bibitem{c4} 	Marler, R. T., Arora, J. S.: Survey of Multi-Objective Optimization Methods for Engineering. In: Struct. Multidisciplinary Optimization, \textbf{26}, pp. 369--395 (2004)
\bibitem{c5} 	Figueira,J., Greco, S., Ehrgott, M. (eds.): Multiple Criteria Decision Analysis: State of the Art Surveys. In: New York (NY), Springer (2005)
\bibitem{c6} 	Eichfelder, G.: Scalarizations for Adaptively Solving Multi-Objective Optimization Problems. In: Comput. Optim. Appl., \textbf{44}, pp. 249--273 (2009)
\bibitem{c7} 	Strongin, R., Sergeyev, Ya.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. In: Kluwer Academic Publishers, Dordrecht (2000, 2nd ed. 2013, 3rd ed. 2014)
\bibitem{c8} 	Sergeyev, Y.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting Space-Filling Curves. In: Springer (2013)
\bibitem{c9} 	Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. In: Princeton University Press (2016)
\bibitem{c10}	Locatelli, M., Schoen, F.: Global Optimization: Theory, Algorithms, and Applications. In: SIAM (2013)
\bibitem{c11}	Sergeyev, Y.D., Grishagin, V.A.: Parallel Asynchronous Global Search and the Nested Optimization Scheme. In: J. Comput. Anal. Appl., \textbf{3}(2), pp. 123--145 (2001)
\bibitem{c12}	Marler, R. T., Arora, J. S.: Multi-Objective Optimization: Concepts and Methods for Engineering. In: VDM Verlag (2009)
\bibitem{c13}	Hillermeier, C., Jahn, J.: Multiobjective Optimization: Survey of Methods and Industrial Applications. In: Surv. Math. Ind., \textbf{11}, pp. 1--42 (2005)
\bibitem{c14}	Gergel, V., Sidorov, S.: A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems. In: LNCS, Springer-Verlag, Berlin Heidelberg, \textbf{9251}, pp. 505--515 (2015)
\bibitem{c15}	Gergel, V.: An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems. In: 2nd International Conference on Mathematics and Com-puters in Sciences and in Industry, pp. 13--18 (2015)
\bibitem{c16}	Gergel, V., Kozinov, E.: Efficient Multicriteria Optimization Based on Intensive Reuse of Search Information. In: Journal of Global Optimization, \textbf{71}(1), pp. 73--90 (2018) 
\bibitem{c17}	Evtushenko, Yu.G., Posypkin, M.A.: A Deterministic Algorithm for Global Multi-Objective Optimization. In: Optimization Methods and Software, \textbf{29}(5), pp. 1005--1019 (2014)
\bibitem{c18}	{\v Z}ilinskas, A., {\v Z}ilinskas, J.: Adaptation of a One-Step Worst-Case Optimal Univariate Algorithm of Bi-Objective Lipschitz Optimization to Multidimensional Problems. In: Commun Nonlinear Sci Numer Simulat, \textbf{21}, pp. 89--98 (2015)
\bibitem{c19}	Pardalos, P.M., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-Convex Multi-Objective Optimization. In: Springer (2017)
\bibitem{c20}	Gergel, V., Kozinov, E.: Accelerating Parallel Multicriterial Optimization Methods Based on Intensive Using of Search Information. In: Procedia Computer Science, \textbf{108}, pp. 1463--1472 (2017) 
\bibitem{c21}	Bleuler, S., Laumanns, M., Thiele, L., Zitzler, E.: PISA-A Platform and Programming Language Independent Interface for Search Algorithms. Evolutionary Multi-Criterion Optimization. In: LNCS, Springer-Verlag, Berlin Heidelberg, \textbf{2632}, pp. 494--508 (2003)
\bibitem{c22}	Chiandussi, G., Codegone, M., Ferrero, S., Varesio, F.E.: Comparison of Multi-Objective Optimization Methodologies for Engineering Applications. In: Comput. Math. Appl., \textbf{63}(5), pp. 912--942 (2012)
\bibitem{c23}	Gaviano, M., Kvasov, D.E., Lera, D., Sergeyev, Ya.D.: Software for Generation of Classes of Test Functions with Known Local and Global Minima for Global Optimization. In: ACM Transactions on Mathematical Software, \textbf{29}(4), pp. 469--480 (2003)

\bibitem{c50}	Borisenko, A., Gorlatch, S.: Parallelizing Metaheuristics for Optimal Design of Multiproduct Batch Plants on GPU. In: LNCS, Springer, Cham, \textbf{10421}, pp. 405--417 (2017)

\end{thebibliography}

\end{document}
% end of file template.tex

### 2018-12-11 2019/PACT_MCO/2019_pavt.tex

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{svproc}
%\usepackage{marvosym}
%
% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{bbding}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\usepackage{hyperref}
\def\UrlFont{\rmfamily}
\providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}

\def\orcidID#1{\unskip$^{[#1]}$}
\def\letter{$^{\textrm{(\Letter)}}$}

\begin{document}
\mainmatter              % start of a contribution
%
\title{Comparative Analysis of Parallel Computational Schemes for 
Solving the Time-consuming Decision Making Problems}
%
\titlerunning{Solving the Time-consuming Decision Making Problems}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Victor Gergel \orcidID{0000-0002-4013-2329} (\Envelope) \and
Evgeny Kozinov \orcidID{0000-0001-6776-0096}}
%
\authorrunning{Victor Gergel et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Victor Gergel and Evgeny Kozinov}
%
\institute{Lobachevsky State University of Nizhni Novgorod \\
Nizhni Novgorod, Russia \\
\email{gergel@unn.ru}, \email{evgeny.kozinov@itmm.unn.ru}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
In the present paper, an efficient approach to the parallel solving of the complex multicriterial optimization problems using the heterogeneous computing systems is considered. The complexity of these problems can be very high as the optimized criteria can be multiextremal and the computing of the criteria values can be time-consuming.  In the framework of the proposed approach the multicriterial optimization problem is reduced to solving a series of global optimization problems by means of the convolution of the partial criteria with different sets of parameters. To solve the series of the global optimization problems, the efficient information-statistical global search method is applied.  The parallel computations are implemented by solving several global optimization problems simultaneously. The comparative analysis of various methods for parallel computations and numerous numerical experiments are presented in the paper.
\keywords{Decision making $\cdot$ Multicriteria global optimization $\cdot$ Parallel calculations $\cdot$ Dimensionality reduction $\cdot$ Search information $\cdot$ Numerical experiments}
\end{abstract}
%
\section{ Introduction}
\label{sec:01}
Decision making is inherent to almost all fields of human activities. A variety of the mathematical formulations have been proposed for the formal description of the decision making problems. Among these ones, there are many classes of optimization problems including unconstrained optimization, nonlinear programming, global optimization, etc. In the most complex situations of decision making, the multicriterial optimization (MCO) problem statements are used widely. An opportunity to set several criteria is a distinctive property of the MCO problems that allows more precise formulating the requirements to the optimality of chosen decisions. The MCO topic is a field of intensive studies - see, for example, the monographs \cite{x1,x2,x3,x4,x5,x6} and the reviews of the scientific and practical results in \cite{x7,x8,x9,x10}.

The possibility of contradictions between the partial criteria of efficiency that makes it impossible to achieve the optimal (the best) values for all partial criteria simultaneously is an important feature of the multicriterial optimization problems. As a result, the finding of some compromised (efficient, non-dominated) decisions, in which the achieved values with respect to some partial criteria satisfy given requirements to necessary level of efficiency, is taken as a solution of an MCO problem usually.

The present work is devoted to solving the MCO problems, which are used for the description the decision making problems in the design of complex objects and systems. In such applications, the partial criteria can take a complex multiextremal form, and computing the values of the criteria and constraints may require a large amount of computations.

Besides, within the framework of considered approach, an opportunity of correcting the MCO problem statements when changing the perceptions of the requirements to the optimality of the chosen decisions is allowed. Thus, in the case of redundancy of the partial criteria set, an opportunity to transform some criteria into the constraints is allowed. Or, otherwise, if the set of feasible variants is insufficient, some constraints can be transformed into the criteria, etc. Such an opportunity to change the MCO problem statements is an additional source of increased computational costs of the search for the optimal decisions.

In the present paper, the results of the performed investigations on the generalization of the decision making problem statements \cite{x11,x12} and on the development of the highly efficient global optimization methods utilizing the whole search information obtained in the course of computations \cite{x13,x14,x15,x16,x17} are presented.

Further structure of the paper is as follows. In Section \ref{sec:02}, the multicriterial optimization problem statement is given. In Section \ref{sec:03}, the reduction of the multicriterial problems to the scalar optimization ones by means of the minimax convolution of partial criteria and the dimensionality reduction with the use of Peano space-filling curves are considered. In Section \ref{sec:04}, the possibility of enhancing the efficiency of computations based on the reuse of the search information is substantiated. In Section \ref{sec:05}, the general organization scheme of parallel computations allowing the maximum use of the computational potential of modern supercomputer systems is described. Section \ref{sec:06} presents the results of numerical experiments supporting the proposed approach to be promising. In conclusion, the obtained results are discussed and possible main directions of further investigations are outlined.

\section{Multiple multicriterial optimization problem statement}
\label{sec:02}
For the formal description of the process of search for the efficient variants in complex decision making problems, the following generalized multilevel model is proposed.

1. At the highest level, within the framework of proposed model, a decision-making problem (DMP) is defined, for which a choice of the optimal values of the parameters should be done according to the available requirements to optimality. In the most general form, DMP can be defined with the use of a vector function of characteristics 
\begin{equation}
\label{eq:01}
w(y)=(w_1(y), w_2(y), \dots, w_M(y)), y \in D,
\end{equation}
where $y=(y_1,y_2, \dots ,y_N)$ is the vector of constructive parameters, and $D \in R^N$ is the domain of feasible values, which usually is a $N$-dimensional hyperinterval
\begin{equation}
\label{eq:02}
D=\{ y\in R^N : a_i \leq y_i \leq b_i,1 \leq i \leq N \},
\end{equation}
for given vectors $a$ and $b$.

The values of characteristics $w(y)$ are supposed to be non-negative, and the decreasing of these ones corresponds to the increasing of the efficiency of the chosen variants. Also, it is supposed that the characteristics $w_j (y)$, $1 \leq j \leq M$ can be multiextremal and the computing of the values of these ones can require a large amount of computations. Besides, within the framework of considered approach, the characteristics $w_j(y)$, $1 \leq j \leq M$ are supposed to satisfy the Lipschitz condition 
\begin{equation}
\label{eq:03}
|w_j(y_1) - w_j(y_2)| \leq L_j \|y_1-y_2\|, 1 \leq j \leq M,
\end{equation}
where $L_j$ is the Lipschitz constant for the characteristic $w_j(y)$, $1 \leq j \leq M$, and  $\|*\|$ denotes the Euclidean norm in $R^N$. It is important to note that the fulfillment of the Lipschitz condition corresponds to the practical applications -- at small variations of the parameters $y \in D$ the corresponding variations of the values of characteristics $w_j(y)$, $1 \leq j \leq M$ are limited as a rule.

In general, the model of DMP is defined by the following set of elements 
\begin{equation}
\label{eq:04}
S=<y,w(y),D,a,b>.
\end{equation}

This model is supposed to remain the same and doesn\'t change in the course of computations.

2. The requirements to the optimality of the chosen variants $y \in D$ of the DMP can be defined in the following way. First of all, the characteristics $w_j(y)$, $1 \leq j \leq M$ should be selected, for which the achievement of the minimum possible values is necessary. Defining a set of indices $F=\{ i_1,i_2, \dots ,i_s \}$ of such characteristics defines the vector criterion of efficiency 
\begin{equation}
\label{eq:05}
f(y)=(w_{i_1}(y),w_{i_2}(y),\dots,w_{i_s}(y)).
\end{equation}

For the characteristics $w_j(y)$, $1 \leq j \leq M$ not included into the vector criterion of efficiency with the set of indices $G=\{ j_1,j_2, \dots ,j_m\}$ the sufficient levels of efficiency should be defined by a tolerance vector $q=(q_1,q_2,\dots,q_m)$. The availability of tolerances allows defining a vector function of constraints 
\begin{equation}
\label{eq:06}
g(y)= (g_1(y),g_2(y),\dots,g_m (y)), g_l (y) = w_{j_l}(y) - q_l, j_l \in G, 1 \leq l \leq m.
\end{equation}

The constraints $g(y)$ define the feasible search domain 
\begin{equation}
\label{eq:07}
Q=\{ y\in D:g(y) \leq 0 \}.
\end{equation}

The criteria of efficiency and constraints formulated in such a way allow formulating a multicriterial optimization (MCO) problem 
\begin{equation}
\label{eq:08}
f(y) \to min, y \in Q,
\end{equation}
defined on the basis of the model $S$ from (\ref{eq:04}) with the use of the elements listed above 
\begin{equation}
\label{eq:09}
P=<S,F,G,q>.
\end{equation}

The scheme proposed above covers many existing optimization problem statements. When $s=1$ and $m=0$ the general statement becomes the global optimization problem. When $s=1$ and $m>0$ the general statement defines a nonlinear programming problem. When $s>1$ and $m>0$ the general statement is reduced to a constrained multicriterial problem.

3.	The use of the multicriterial optimization problem statements reduces the complexity of the formal description of the decision making problems -- it becomes possible to define several partial criteria instead of necessity to define a unified ''global'' criterion of efficiency in the MCO problems. No doubt, such an approach is a simpler task for the decision maker determining the requirements to the optimality of the decisions made. It is worth noting also that the multicriterial definition of the efficiency criteria corresponds to the practice of decision making in various fields of applications.

The scheme of the MCO problem statement considered above has been proposed first in \cite{x11} and was used widely in solving many applied decision making problems. At the same time, the results of the practical application of the approach have demonstrated that the formulating of a single MCO problem statement might also appear to be difficult when altering the ideas of the necessary requirements to the optimality. Thus, in the case of redundancy of the partial criteria set, the transformation of some criteria into the constraints may appear to be appropriate. Or, vise versa, if a small feasible domain $Q$ is obtained, the loosening of the tolerances $q$ or transforming some constraints into the criteria may be required. As a result, an opportunity of simultaneous formulating several MCO problems 
\begin{equation}
\label{eq:10}
\mathbb{P}_t=\{P_1,P_2,\dots,P_t\}
\end{equation}
will be allowed  within the framework of the extended model of the optimal choice as further development of the scheme (\ref{eq:01})-(\ref{eq:09}). It is worth noting that the set of problems $\mathbb{P}$ can change in the course of computations by means of adding new problems 
\begin{equation}
\label{eq:11}
\mathbb{P}'=\mathbb{P}+/-\mathbb{P}.
\end{equation}
or removing already existing ones. Solving the problems from the set $\mathbb{P}$ can be performed sequentially or simultaneously in the time-sharing mode or in parallel if several computing devices are available. No doubt, the simultaneous method of solving the problems is more preferable -- the results obtained in the course of computations allow adjusting current set of problems $\mathbb{P}$ promptly. In the case of utilizing the parallel computations, the total time of problem solving can be reduced essentially.

In general, the proposed model of the process of search for the optimal decisions (\ref{eq:01})-(\ref{eq:11}) defines a new class of the optimization problems -- the multiple multicriterial global optimization (MMGO) problems.

\section{Successive reduction of the multicriterial global optimization problems }
\label{sec:03}

As it has been already mentioned above, the partial criteria of efficiency in the MCO problems are contradictory usually. This feature of the MCO problems means that the obtaining of the optimal (the best) values with respect to all partial criteria simultaneously cannot be ensured. In such situation, the finding of particular compromised (effective, non-dominated) decisions, in which the achieved values with respect to particular partial criteria cannot be improved without simultaneous worsening with respect to some other criteria is understood as the solution of a MCO problem usually. In the limiting case, the finding of all effective (Pareto-optimal) decisions $PD(Q) \subset Q$. can be required.

Due to a high relevance, a large enough quantity of methods of solving the MCO problems have been developed and used widely -- see, for example, \cite{x2,x3,x4,x5,x6,x7,x8,x9,x10}. A part of these algorithms ensures the obtaining the numerical estimates of the whole Pareto set $PD(Q)$ \cite{x3,x26,x27,x28}. Along with the usefulness of such an approach (all effective decisions of the MCO problem are found), the use of these methods appears to be difficult often because of high computational complexity of obtaining the estimates of the Pareto set. Besides, the estimate of the whole set $PD(Q)$ may appear tо be redundant when to solve the MCO problem it is enough to obtain several particular effective decisions (that takes place in the practical applications often). As a result, more widely used approach to solving the MCO problems is based on the scalarization of the vector criterion into some general scalar criterion of efficiency, the optimization of which can be performed with the use of already existing optimization methods. Among these methods, for example, there are weighted sum method, compromise programming method, weighted min-max method, goal programming, and many other algorithms -- see, for example, \cite{x2,x3,x4,x5,x6}.

The general approach used in the present work consists in the reduction of solving a MMGO problem to the solving of a series of single-criterion global optimization problems \footnote{Due to initial assumptions on possible multiextremality of the characteristics $w_j(y), 1 \leq j \leq M$ from (\ref{eq:01})}: 
\begin{equation}
\label{eq:12}
\min {\varphi(y)}=F(\lambda,y), g(y) \leq 0, y \in D, 
\end{equation}
where $g(y)$ is the vector function of constraints from (\ref{eq:06}), and $F(\lambda,y)$ is the minimized convolution of the partial criteria $f_i(y(x))$, $1 \leq i \leq s$ of the MCO problem (\ref{eq:05})
\begin{equation}
\label{eq:13}
F(\lambda,y)=\max{⁡(\lambda_i f_i (y), 1 \leq i \leq s)} 
\end{equation}
using the vector of the weighting coefficients 
\begin{equation}
\label{eq:14}
\lambda=(\lambda_1, \lambda_2, \dots , \lambda_s )\in \Lambda \subset R^s: \sum_{i=1}^s{\lambda_i=1}, \lambda_i \geq 0, 1 \leq i \leq s.
\end{equation}

In the developed approach, it is proposed to perform one more step of conversion of the problems $F(\lambda,y)$ from (\ref{eq:12}) being solved -- to perform the dimensionality reduction using the Peano space-filling curves (evolvents) $y(x)$ providing an unambiguous mapping of the interval $[0,1]$ onto an $N$-dimensional hypercube D \cite{x18,x20}. As a result of such reduction, the multidimensional global optimization problem (\ref{eq:12}) is reduced to the one-dimensional problem:
\begin{equation}
\label{eq:15}
F(\lambda,y(x^*)) = \min{\{F(\lambda,y(x)): g(y(x)) \leq 0, x \in [0,1]\}}.
\end{equation}

The considered dimensionality reduction scheme superimposes the multidimensional problem with the Lipschitzian minimized function $F(\lambda,y)$ to a one-dimensional problem, in which the reduced function $F(\lambda,y(x))$ satisfies the uniform H\"{o}lder condition i.e.
\begin{equation}
\label{eq:16}
|F(\lambda,y'(x)) - F(\lambda, y''(x))| \leq H \|x' - x''\|, x', x'' \in [0,1], 
\end{equation}
where the constant $H$ is defined by the relation $H=2L\sqrt{N+3}$, $N$ is the dimensionality of the optimization problem from (\ref{eq:01}), and $L$ is the Lipschitz constant of the function $F(\lambda,y)$ from (\ref{eq:12}). The dimensionality reduction allows applying many well known highly efficient one-dimensional global optimization algorithms for solving the problems (\ref{eq:12}) (after performing the necessary generalization) -- see, for example, \cite{x18,x20,x29,x30,x31,x32,x33,x34,x35}.

In order to obtain several effective decisions (or to estimate the whole Pareto domain) the problem (\ref{eq:12}) should be solved several times for corresponding sets of the values of the elements of the weighting vector $\lambda \in \Lambda$. In this case, the set of MCO problems $\mathbb{P}$ from (\ref{eq:10}) necessary for solving the initial decision making problem is transformed into a wider set of scalar optimization problems (\ref{eq:12})
\begin{equation}
\label{eq:17}
\mathbb{F}_T=\{F_i (\lambda_i,y): \lambda_i \in \Lambda, 1 \leq i \leq T \},
\end{equation}
where several problems (\ref{eq:12}) with different values of the coefficients $\lambda \in \Lambda$ can correspond to each problem $P \in \mathbb{P}$.

\section{Improvement the efficiency of computations based on the reuse 
of the search information} 
\label{sec:04}
As it has been already mentioned above, the solving of the MCO problems $P$ from (\ref{eq:10}) and corresponding global optimization problems $F$ from (\ref{eq:15}) can require a large amount of computations even when solving a single particular problem. In the case if a series of problems from the sets $\mathbb{P}$ and $\mathbb{F}$ should be solved, the necessary amount of computations could appear to be even larger. And, therefore, overcoming the computational complexity of the decision making problems of the considered class is a necessary condition of opportunity of a practical application of the proposed approach. An intensive use of the whole search information obtained in the course of computations can be a promising approach to solving this problem.

The numerical solving of the global optimization problems is reduced usually to a sequential computing the values of characteristics $w(y)$ at the points $y^i$, $0\leq i \leq k$ of the search domain $D$ \cite{x18,x22,x23,x24,x25}. The data obtained as a result of computations can be represented in a form of matrix of search information (MSI)
\begin{equation}
\label{eq:18}
\Omega_k=\{(y^i,w^i=w(y^i ) )^T:1 \leq i \leq k\}.
\end{equation}

As a result of scalarization of the vector criterion (\ref{eq:12}) and application of the dimensionality reduction (\ref{eq:15}), $\Omega_k$ from (\ref{eq:18}) can be transformed into the form of the matrix of the search state (MSS):
\begin{equation}
\label{eq:19}
A_k=\{(x_i,z_i,g_i,l_i )^T:1 \leq i \leq k\},
\end{equation}
where
\begin{itemize}
	\item $x_i$, $1 \leq i \leq k$ are the reduced points of the executed global search iterations where the criteria values have been computed; 
	\item $z_i$, $g_i$, $1 \leq i \leq k$ are the values of the scalar criterion and of the constraints at the points $x_i$, $1 \leq i \leq k$ for current optimization problem $F(\lambda,y(x))$ from (\ref{eq:15}) being solved i. e.
\begin{equation}
z_i = \varphi(x_i) =F(\lambda,y(x_i)),  g_i=g(y(x_i )),  1\leq i \leq k;
\end{equation}
	\item $l_i$, $1 \leq i \leq k$ are the indices of the global search iterations, at which the points $x_i$, $1 \leq i \leq k$ have been generated, these indices are used to storing the correspondence between the reduced points and the multidimensional ones at the executed iterations i. e.
\begin{equation}
y^j = y(x_i), j = l_i, 1 \leq i \leq k.
\end{equation}

\end{itemize}

The matrix of the search state $A_k$ contains the search information transformed to current scalar reduced problem (\ref{eq:15}) being solved. Besides, the search information in the MSS is arranged according to the values of coordinates of the points $x_i$, $1 \leq i \leq k$ for more efficient execution of the global search algorithms; the arranged representation of the points is reflected by the use of the lower index i. e.
\begin{equation}
x_1 \leq x_2 \leq \dots \leq x_k.
\end{equation}

The representation of the search information obtained in the course of computations in the form of  the matrices $\Omega_k$ and $A_k$ makes the basis for the development of efficient optimization search procedures. The availability of such information allows performing an adaptive choice of the points for executing the global search iterations taking into account the results of all computations completed earlier 
\begin{equation}
\label{eq:20}
y^{k+1}=Y(\Omega_k),k=1,2,\dots.
\end{equation}
where $Y$ is the rule for computing the points $y^i$, $0 \leq i \leq k$ of the applied optimization algorithm. Such an adaptive choice of the points of executed search iterations can accelerate the finding of the effective decisions. In the case of the global optimization problems, the accumulation of the whole search information and the use of the rules of the kind (\ref{eq:20}) is just mandatory. Any reduction of the search information in determining the points to be executed would result in executing the excess global search iterations.

It should be noted that the availability of MSS $\Omega_k$ from (\ref{eq:18}) allows reducing the results of all preceding computations $z_i$,  $1 \leq i \leq k$ stored in the MSS $A_k$ to the values of current optimization problem $F(\lambda,y(x))$ from (\ref{eq:15}) being solved for any new parameter of the MCO problem statement $P$ from (\ref{eq:09}) and for any new value of the convolution coefficients $\lambda$ from (\ref{eq:14}) on the basis of the values of characteristics $w^i$, $1 \leq i \leq k$ computed earlier and stored in $\Omega_k$ without any repeated time-consuming computation of the values of $w(y)$ from (\ref{eq:01}) i.e.
\begin{equation}
\label{eq:21}
w_i \xrightarrow{\lambda,P} (z_i,g_i ), 1 \leq i \leq k,\forall \lambda \in \Lambda, P \in \mathbb{P}.
\end{equation}

Therefore, all the search information can be employed for continuing the computations in full amount. An efficient use of this information becomes an important requirement in the selection of the methods for solving the problems $F(\lambda,y(x))$ from (\ref{eq:15}). The reuse of the search information can provide a continuous decreasing of the amount of computations for solving every next optimization problem up to execution of just few iterations only to find the next effective decision.

\section{Parallel computations in the multiple multicriterial global optimization problems} 
\label{sec:05}

The sequential solving the set of problems $\mathbb{F}$ from (\ref{eq:17}) to find several globally optimal decisions increases the amount of computations required to solve the decision making problems essentially. The parallel solving of the problems $F(\lambda,y(x))$ from the set $\mathbb{F}$ is a promising and simple method of accelerating the performed computations. At high enough number of the computer nodes (processors), the time of solving the whole set of problems $\mathbb{F}$ would be determined by the computation time for the problem $F(\lambda,y(x))$, the solving of which takes the longest time. Such an approach can be implemented quite easily. However, it doesn\'t solve the problem of computational complexity when it is necessary to expand the set of problems $\mathbb{F}$ in the course of computations.

The method of organizing the parallel computations considered above can be developed further by utilizing the information compatibility of the problems from the set $\mathbb{F}$ -- as has been noted in Section {\ref{sec:04}}, the values of characteristics $w^i$, $1 \leq i \leq k$ computed when solving any problem $F_1(\lambda,y(x))$ can be transformed to the values of any other problem $F_2(\lambda,y(x))$. This result allows proposing the following general scheme of parallel computations for simultaneous solving the problems from the set $\mathbb{F}$.

\textbf{1. The distribution of problems.} Prior to the start of computations, the problems of the set $\mathbb{F}$ from (\ref{eq:17}) should be distributed among the computational nodes of a multiprocessor system. This distribution can be quite diverse -- for solving a particular problem from the set $\mathbb{F}$, various numbers of computational cores $(q, q \geq 1)$ and of the computational nodes $(p, p \geq 1)$ can be allocated. An opportunity to use several computational nodes $p>1$ for solving the same problem is provided by means of application of multiple evolvents $y_i (x)$, $1\leq i\leq p$ -- see \cite{x13,x14}. If when solving a particular problem from the set $\mathbb{F}$ it is planned not to employ the processor cores of a computational node in full, in this case several optimization problems can be being solved using the same node subject to the number of processors available on the computational node and to the number of computational cores in each processor.

\textbf{2. The choice of the optimization algorithms.} The proposed scheme of the parallel computations is a general one -- various methods of multiextremal optimization can be applied for solving the problems on each computational node -- see, for example, \cite{x18,x20,x29,x30,x31,x32,x33,x34,x35}. The main requirement to the selection of the algorithms -- the methods should use the search information $\Omega
_k$ and $A_k$ to increase the efficiency of global search. And, as it has been already noted above in Section \ref{sec:03}, the global optimization algorithms should be generalized in order to provide solving the reduced one-dimensional global optimization problems of the kind $F(\lambda,y(x))$ from the set $\mathbb{F}$. The possibility of application of the proposed approach was substantiated using the example of the efficient global search algorithms developed within the framework of the information-statistical theory of the multiextremal optimization -- see \cite{x12,x13,x14,x15,x16,x17,x18}.

\textbf{3. The execution of the global search iterations.} The execution of the global search iterations for each problem $F(\lambda,y(x))$ from the set $\mathbb{F}$ being solved is performed in parallel. The execution of every iteration on each computational node includes performing the following operations:
\begin{enumerate}
	\item The choice of $q$, $q \geq 1$ points $x^i$, $1\leq i\leq q$ of the next global search iteration is performed according to the rule (\ref{eq:20}) of the optimization method; the choice of the points is performed taking into account the available search information $\Omega_k$ and $A_k$; the number of generated points $q$, $q\geq 1$ is determined by the number of computational cores employed for solving the problem $F(\lambda,y(x))$;
	\item For each chosen one-dimensional point $x^i \in [0,1]$, $1 \leq i \leq q$ of current global search iteration, the multidimensional image $y^i \in D$, $1 \leq i \leq q$ is determined according to the mapping $y(x)$. Then, each computed image $y^i \in D$, $1 \leq i \leq q$ is transmitted to all employed computational nodes to avoid the multiple choice of the same points in the domain $D$ when solving the problems from the set $\mathbb{F}$;
	\item The values of characteristics $w(y)$ from (\ref{eq:01}) are computed at all points $y^i \in D$, $1 \leq i \leq q$; for each particular point $y^i \in D$, $1 \leq i \leq q$, these computations are performed in parallel using different computational cores; the computed values of characteristics $w(y)$ are transmitted to all employed computational nodes to include the obtained data into the search information $\Omega_k$ and $A_k$.
\end{enumerate}

\textbf{4. Updating the search information.} Prior to starting the next global search iteration, the availability of data transmitted from other computational units (processors or cores) is checked; the received data should be included into the search information.

According to the presented computational scheme of parallel computations, each computational unit would contain the identical copies of the matrix of search information $\Omega_k$ from (\ref{eq:18}); the matrices $A_k$ from (\ref{eq:19}) would contain different sets of the global search points $x_i$, $1\leq i\leq k$ because of the use of different evolvents $y(x)$ for the dimensionality reduction and of different values of the scalar criterion and constraints $z_i$, $g_i$, $1\leq i\leq k$ corresponding to the problems $F(\lambda,y(x))$ from the set $\mathbb{F}$ being solved.

Within the framework of such computational scheme, it becomes possible to expand the set of problems being solved $\mathbb{F}$ at any moment of computations -- in order to solve a new problem $F(\lambda,y(x))$, it is enough to allocate an additional computational unit, to copy the set $\Omega_k$, and to create the corresponding matrix $A_k$.

It is clear that the increase of the efficiency of the global search due to the use of the sets of the search information $\Omega_k$ and $A_k$ depends on the optimization algorithms employed. The analysis of the elements of the scheme of organization of the parallel computations considered above in application to the information-statistical global optimization algorithms has been performed in \cite{x12,x13,x14,x15,x16,x17}. The results of numerical experiments carried out for complete experimental evaluation of the efficiency of the proposed scheme to choose the best parameters of the use of the computational resources of the high-performance supercomputer systems are presented below in Section \ref{sec:06}.

\section{The results of numerical experiments }
\label{sec:06}

The numerical experiments have been carried out using ''Lobachevsky'' supercomputer at University of Nizhni Novgorod (operation system -- CentOS 6.4, management system -- SLURM). Each supercomputer node had 2 Intel Sandy Bridge E5-2660 processors (2.2 GHz, 64 Gb RAM). Each processor included 8 cores (i.e. total 16 CPU cores were available at each computer node). In order to generate the executable program code, Intel C++ 17.0 compiler was used.

When performing the numerical experiments, according to the general scheme of parallel computations considered in Section \ref{sec:05}, the efficient algorithms of multicriterial multiextremal optimization \cite{x12,x13,x14,x15,x16,x17} developed within the framework of the information-statistical theory of global optimization \cite{x18} were used. The efficiency of these algorithms has been proven by many numerical experiments.

In the present work, the results of a series of the numerical experiments performed in order to choose the best parameters of the parallel computations (the number of employed computational cores and computer nodes, the number of the multicriterial optimization subproblems solved simultaneously, the number of evolvents used for the dimensionality reduction, etc.) are presented. Each experiment included the solving of 30 MCO problems. Each problem was a six-dimensional and five-criterial one i.e. $N=6$, $s=5$. The multiextremal functions defined by the GKLS generator \cite{x36} were used as the criteria. This generator allows generating the multiextremal optimization problems with the properties known a priori: the number of local minima, the sizes of their attractors, the global minimum point, the objective function value at this point, etc. 
%Some example contour plots of the functions obtained by the GKLS generator are presented in Fig. \ref{fig:01}. 
%\begin{figure}[htbp]
%\centerline{\includegraphics[width=0.7\textwidth]{fig1.png}}
%\caption{Contour plots of three multiextremal functions from the GKLS family of test optimization problems}
%\label{fig:01}
%\end{figure}

The following values of parameters were used in the numerical experiments. When solving the MCO problems, the condition of stopping the computations according to predefined accuracy $\varepsilon=0.05$ was used. For constructing the estimates of the Pareto domain $PDA(f,D)$, 100 subproblems of the family $\mathbb{F}$ from (\ref{eq:17}) have been solved for various values of the coefficients $\lambda \in \Lambda$. In the performed experiments, 10 computational nodes of the supercomputer were used, each computational node included two 8-core processors (i.e. total 160 computational cores were employed for performing the computations in each experiment).

In the beginning of the performed series of experiments, the parallel computations have been conducted using a single supercomputer node only (two processors, 16 computational cores using the shared memory). The results of these experiments are presented in Table \ref{tab:02}.

%Table 2. Evaluation of the efficiency of using a single supercomputer node when solving 
%a series of 30 six-dimensional five-criterial MCO problems
%Cores	Search information	Iterations	S1	S2
%1	not used	26 813 722.7	1.0	
%1	used 	9 103 069.6	2.9	1.0
%8	used 	1 291 720.0	20.8	7.0
%16	used 	609 169.5	44.0	14.9
\begin{table}[htbp]
\centering
\caption{Evaluation of the efficiency of using a single supercomputer node when solving a series of 30 six-dimensional five-criterial MCO problems}
\label{tab:02}
\begin{tabular}{ccccc}
\hline
Cores & Search information & Iterations   & S1   & S2   \\ \hline
1     & not used           & 26 813 722.7 & 1.0  &      \\
1     & used               & 9 103 069.6  & 2.9  & 1.0  \\
8     & used               & 1 291 720.0  & 20.8 & 7.0  \\
16    & used               & 609 169.5    & 44.0 & 14.9 \\ \hline
\end{tabular}
\end{table}

In Table \ref{tab:02}, the number of computational cores employed when carrying out the experiments is presented in the \textit{Cores} column. The \textit{Search information} column contains the information on the reuse of the search information when solving the series of subproblems of the family $\mathbb{F}$ from (\ref{eq:17}) for different values of the coefficients $\lambda \in \Lambda$. The \textit{Iterations} column presents the average number of the global search iterations executed when solving particular subproblems of the MCO problem. The \textit{S1} column shows the overall speedup achieved by means of performing the parallel computations using the corresponding quantity of the computational cores. The \textit{S2} column contains the speedup of computations with respect to the serial optimization algorithm, in which the search information is used.

As follows from the results presented in Table \ref{tab:02}, when using even a single computational core, one can achieve a speedup of computations almost by a factor of three by means of the reuse of the search information obtained in the course of optimization. The overall speedup achieved in the parallel computations makes 44 times using 16 cores of a single supercomputing node.

Further, in all experiments 10 supercomputer nodes were used (i. e. total 160 computational cores were employed). The parameters of the experiments were: the number of subproblems of the family $\mathbb{F}$ from (\ref{eq:17}) being solved simultaneously, the number of computational nodes employed for solving particular MCO subproblems, and the number of test problems (generated by the GKLS generator) being solved simultaneously (total 30 six-dimensional five-criterial test MCO problems were solved in each experiment). The results of the numerical experiments are presented in Table \ref{tab:03}.
%
%Table 3. Evaluation of efficiency of using ten supercomputing nodes when solving a series 
%of 30 six-dimensional five-criterial MCO problems
%Nodes	NumMCO	SubProblems	SubNodes	Iterations	S1	Scaling
%1	1	1	1	609 169.5	44.0	1.0
%10	1	10	1	75 045.8	357.3	8.1
%10	1	5	2	77 643.6	345.3	7.8
%10	1	1	10	108 700.0	246.7	5.6
%10	2	5	1	76 819.0	349.1	7.9
%10	2	1	5	81 132.8	330.5	7.5
%10	5	2	1	66 618.6	402.5	9.1
%10	5	1	2	77 995.0	343.8	7.8
%10	10	1	1	73 999.7	362.3	8.2
\begin{table}[htbp]
\centering
\caption{Evaluation of efficiency of using ten supercomputing nodes when solving a series of 30 six-dimensional five-criterial MCO problems}
\label{tab:03}
\begin{tabular}{ccccccc}
\hline
Nodes       & NumMCO     & SubProblems & SubNodes   & Iterations        & S1             & Scaling      \\ \hline
1           & 1          & 1           & 1          & 609 169.5         & 44.0           & 1.0          \\
10          & 1          & 10          & 1          & 75 045.8          & 357.3          & 8.1          \\
10          & 1          & 5           & 2          & 77 643.6          & 345.3          & 7.8          \\
10          & 1          & 1           & 10         & 108 700.0         & 246.7          & 5.6          \\
10          & 2          & 5           & 1          & 76 819.0          & 349.1          & 7.9          \\
10          & 2          & 1           & 5          & 81 132.8          & 330.5          & 7.5          \\
\textbf{10} & \textbf{5} & \textbf{2}  & \textbf{1} & \textbf{66 618.6} & \textbf{402.5} & \textbf{9.1} \\
10          & 5          & 1           & 2          & 77 995.0          & 343.8          & 7.8          \\
10          & 10         & 1           & 1          & 73 999.7          & 362.3          & 8.2          \\ \hline
\end{tabular}
\end{table}

The number of employed supercomputing nodes is given in the \textit{Nodes} column. In the \textit{NumMCO} column, the number of test MCO problems (generated by the GKLS generator) solved simultaneously is given. In the \textit{SubProblems} column, the number of subproblems of the family $\mathbb{F}$ from (\ref{eq:17}) solved simultaneously is presented. In the \textit{SubNodes} column, the number of the computational nodes employed for the solving of particular MCO subproblems is presented. The \textit{Iterations} column presents the averaged number of the global search iterations executed when solving particular subproblems of the MCO problem. The \textit{S1} column shows the overall speedup achieved by means of performing the parallel computations with the use of the corresponding quaintly of the supercomputing nodes. The \textit{Scaling} column contains the speedup of the parallel computations with respect to the results obtained when using a single computational node only.

As follows from the results presented in Table \ref{tab:03}, the best speed up in the considered series of experiments has been achieved at the simultaneous solving of five MCO problems of the test class when two subproblems of the family $\mathbb{F}$ from (\ref{eq:17}) are solved in parallel for each problem using a single supercomputing node for each subproblem (the parameters for this experiments are highlighted by the bold color in Table \ref{tab:03}). In this case, the speedup achieved for the parallel computations exceeds 400 (it should be noted that total 10 computational nodes including total 160 cores were used in this experiment).

\section{Conclusion }
\label{sec:07}

In the present paper, an efficient computational scheme of the parallel computations for solving the complex multicriterial optimization problems with the non-convex constraints, in which the optimization criteria can be multiextremal and computing the criteria values can require a large amount of computations is proposed. The proposed approach is based on the reduction of the multicriterial problems to the nonlinear programming ones by the minimax convolution of the partial criteria, the dimensionality reduction using the Peano space-filling curves, and application of the efficient information-statistical global optimization methods using a novel index scheme of accounting for the constraints instead of the penalty functions used commonly.

The developed general computational scheme includes the concurrent parallel computational methods for the computational systems with the shared memory, the distributed parallel computations using the multiple mappings for the dimensionality reduction, and the multilevel nesting of the parallel computations for the high-performance computational systems. In general, the proposed computational scheme ensures the opportunity of an efficient application of the high-performance computational systems with a large number (tens and hundreds thousand) cores/processors for solving the complex global optimization problems. Besides, this general scheme can be used for the organization of the parallel computations for a wide range of algorithms for solving the time-consuming problems of decision making (in particular, for the information-statistical global optimization algorithms).

In order to determine the optimal parameters of the parallel computations, a series of experiments has been carried out. The results have shown the developed approach to allow reducing the computational costs of solving the multicriterial optimization problems with non-convex constraints considerably.

In conclusion, one can note the developed approach to be promising and requires further investigations. First of all, it is necessary to continue carrying out the numerical experiments on solving the multicriterial optimization problems for a larger quantity of partial criteria of efficiency and for larger dimensionality of the optimization problems being solved.

\section*{Acknowledgements}  
This research was supported by the Russian Science Foundation, project No 16-11-10150 ''Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance.''

%
% ---- Bibliography ----
%
\bibliographystyle{spmpsci}
\begin{thebibliography}{6}
%

%\bibitem {czaj:fitz}
%Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid information services
%for distributed resource sharing. In: Proceedings 10th IEEE International Symposium
%on High Performance Distributed Computing, pp. 181--184. IEEE, New York (2001).
%\doi{10.1109/HPDC.2001.945188}
%
%\bibitem {fost:kes}
%Foster, I., Kesselman, C. (eds.): The grid: blueprint for a new computing infrastructure.
%Morgan Kaufmann, San Francisco, CA, USA (1999)
%
%\bibitem {fo:kes:nic:tue}
%Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The physiology of the grid: an open grid services architecture for distributed systems integration. Tech. rep., Global Grid
%Forum (2002)
%
%\bibitem {may:ehr:stein}
%May, P., Ehrlich, H.-C., Steinke, T.: ZIB structure prediction pipeline:
%composing a complex biological workflow through web services.
%In: Nagel, W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006.
%LNCS, vol. 4128, pp. 1148--1158. Springer, Heidelberg (2006).
%\doi{10.1007/11823285\_121}
%
%\bibitem {onlyurl}
%NCBI: National Center for Biotechnology Information. \url{http://www.ncbi.nlm.nih.gov}
%
%\bibitem {smit:wat}
%Smith, T.F., Waterman, M.S.: Identification of common molecular subsequences.
%Journal of Molecular Biology. \textbf{147}(1), 195--197 (1981). \doi{10.1016/0022-2836(81)90087-5}

\bibitem {x1} Parnell, G.S., Driscoll, P.J., Henderson, D.L., editors.: Decision Making in Systems Engineering and Management. Wiley, New Jersey (2nd ed., 2011).
\bibitem {x2} Miettinen K.: Nonlinear Multiobjective Optimization. Springer (1999).
\bibitem {x3} Marler, R. T., Arora, J. S.: Multi-Objective Optimization: Concepts and Methods for Engineering. VDM Verlag (2009).
\bibitem {x4} Ehrgott, M.: Multicriteria Optimization. Springer (2nd ed., 2010).
\bibitem {x5} Collette, Y., Siarry, P.: Multiobjective Optimization: Principles and Case Studies (Decision Engineering). Springer (2011).
\bibitem {x6} Pardalos, P.M., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-Convex Multi-Objective Optimization. Springer (2017). 
\bibitem {x7} Figueira,J., Greco, S., Ehrgott, M., editors.: Multiple criteria decision analysis: State of the art surveys. New York, Springer (2005).
\bibitem {x8} Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications. Surv. Math. Ind. \textbf{11}, 1--42 (2005).
\bibitem {x9} Zavadskas, E. K., Turskis, Z., Kildiene, S.: State of art surveys of overviews on MCDM/MADM methods. Technological and Economic Development of Economy \textbf{20}, 165--179 (2014).
\bibitem {x10}Cho, J.-H., Wang, Ya., Chen, I.-R., Chan, K.S., Swami, A.: A Survey on Modeling and Optimizing Multi-Objective Systems. IEEE Communications Surveys \& Tutorials \textbf{19}(3), 1867--1901 (2017). 
\bibitem {x11}Strongin, R.G., Gergel, V.P. Markin, D.L.:  Multicriterion multiextreme optimization with nonlinear constraints. Lecture Notes in Economics and Mathematical Systems, \textbf{351}, 120--127 (1988).
\bibitem {x12}Gergel, V.P., Kozinov, E.A.: Accelerating multicriterial optimization by the intensive exploitation of accumulated search data. AIP Conference Proceedings \textbf{1776}, 090003 (2016). \doi{10.1063/1.4965367}
\bibitem {x13}Gergel, V.P., Kozinov, E.A.: Efficient multicriterial optimization based on intensive reuse of search information. J. Glob. Optim. \textbf{71}(1), 73--90 (2018). \doi{doi.org/10.1007/s10898-018-0624-3}
\bibitem {x14}Gergel, V.P., Kozinov, E.A.: Accelerating Parallel Multicriterial Optimization Methods Based on Intensive Using of Search Information. Procedia Computer Science, \textbf{108}, 1463--1472 (2017).
\bibitem {x15}Gergel, V., Kozinov, E.: Parallel computing for time-consuming multicriterial optimization problems. LNCS, \textbf{10421}, 446--458 (2017).  \doi{10.1007/978-3-319-62932-2\_43}
\bibitem {x16}Gergel, V., Kozinov, E.: Efficient methods of multicriterial optimization based on the intensive use of search information. Springer Proceedings in Mathematics and Statistics, \textbf{197}, 27--45 (2017). \doi{10.1007/978-3-319-56829-4\_3}
\bibitem {x17}Gergel, V., Kozinov, E.: An approach for parallel solving the multicriterial optimization problems with non-convex constraints. Communications in Computer and Information Science, \textbf{793}, 121--135 (2017). \doi{10.1007/978-3-319-71255-0\_10}
\bibitem {x18}Strongin, R., Sergeyev, Ya.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2nd ed. 2013, 3rd ed. 2014).
\bibitem {x19}Strongin, R., Gergel, V., Grishagin, V., Barkalov, K.: Parallel computations for global optimization problems. Moscow State University Press (2013). (in Russian)
\bibitem {x20}Sergeyev Y.D., Strongin R.G., Lera D.: Introduction to global optimization exploiting space-filling curves. Springer (2013).
\bibitem {x21}Eichfelder, G.: Scalarizations for adaptively solving multi-objective optimization problems. Comput. Optim. Appl. \textbf{44}, 249--273 (2009). 
\bibitem {x22}Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. Princeton University Press (2016).
\bibitem {x23}Locatelli, M., Schoen, F.: Global Optimization: Theory, Algorithms, and Applications. SIAM (2013).
\bibitem {x24}Arora, R.K.: Optimization: Algorithms and Applications. CRC Press (2015).
\bibitem {x25}Bazaraa, M.S., Sherali, H.D., Shetty, C.M.: Nonlinear Programming: Theory and Algorithms, John Wiley and Sons (2006).
\bibitem {x26}Evtushenko, Yu.G., Posypkin, M.A.: A deterministic algorithm for global multi-objective optimization. Optimization Methods and Software, \textbf{29}(5), 1005--1019 (2014).
\bibitem {x27}{\v Z}ilinskas, A., {\v Z}ilinskas, J.: Adaptation of a one-step worst-case optimal univariate algorithm of bi-objective Lipschitz optimization to multidimensional problems. Commun Nonlinear Sci Numer Simulat, \textbf{21}, 89--98 (2015).
\bibitem {x28}Markin, D.L., Strongin, R.G.: Uniform estimates for the set of weakly effective points in multi-extremum multicriterion optimization problems. Computational Mathematics and Mathematical Physics \textbf{33}(2), 171--179 (1993).
\bibitem {x29}Piyavskij, S.A.: An algorithm for finding the absolute extremum of a function. Computational Mathematics and Mathematical Physics \textbf{12}, 57--67 (1972).
\bibitem {x30}Shubert, B.O.: A sequential method seeking the global maximum of a function. SIAM J. Numer. Anal. \textbf{9}, 379--388 (1972).
\bibitem {x31}Breiman, L., Cutler, A.: A deterministic algorithm for global optimization. Math. Program. \textbf{58} (1--3), 179--199 (1993). 
\bibitem {x32}T\"{o}rn, A., {\v Z}ilinskas, A.: Global Optimization, Springer-Verlag, Lecture Notes in Computer Science, vol. \textbf{350} (1989).
\bibitem {x33}Evtushenko, Y.G.: Numerical Optimization Techniques. Translations Series in Mathematics and Engineering, Springer-Verlag, Berlin (1985).
\bibitem {x34}Sergeyev, Y.D.: Global one-dimensional optimization using smooth auxiliary functions. Mathematical Programming \textbf{81}, 127--146 (1998).
\bibitem {x35}Gergel, V.P.: A method of using derivatives in the minimization of multiextremum functions. Computational Mathematics and Mathematical Physics \textbf{36}, 729--742 (1996). 
\bibitem {x36}Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003).

\end{thebibliography}
\end{document}

### 2018-12-28 2019/MOTOR/motor.tex
Blank paper

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{ Generalized nested optimization scheme for global optimization problems
\thanks{This study was supported by the Russian Science Foundation, project 
No.\,16-11-10150.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set 
% an abbreviated paper title here
%
\author{Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
Ilia Lebedev\orcidID{0000-0002-8736-0652}}
%
\authorrunning{K. Barkalov, I. Lebedev}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
\email{konstantin.barkalov@itmm.unn.ru}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{Dimensionality reduction \and Global optimization \and Multiextremal functions.}
\end{abstract}
%
%
%
\section{First Section}
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

%\begin{figure}
%\includegraphics[width=\textwidth]{fig1.eps}
%\caption{A figure caption is always placed below the illustration.
%Please note that short captions are centered, while long ones are
%justified by the macro package automatically.} \label{fig1}
%\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{globalizerSystem}, a
book~\cite{strSergGO}, proceedings without editors~\cite{Jones2009}.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
 \bibliographystyle{splncs04}
 \bibliography{bibliography}
%

%\begin{thebibliography}{8}
%\bibitem{ref_article1}
%Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
%
%\bibitem{ref_lncs1}
%Author, F., Author, S.: Title of a proceedings paper. In: Editor,
%F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
%Springer, Heidelberg (2016). \doi{10.10007/1234567890}
%
%\bibitem{ref_book1}
%Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
%Location (1999)
%
%\bibitem{ref_proc1}
%Author, A.-B.: Contribution title. In: 9th International Proceedings
%on Proceedings, pp. 1--2. Publisher, Location (2010)
%
%\bibitem{ref_url1}
%LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
%Oct 2017
%\end{thebibliography}
\end{document}

### 2019-01-16 2019/WCGO/WCGO.tex
Нулевая версия статьи на WCGO

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}

% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{ Generalized scheme for dimensionality reduction based on nested 
optimization and Peano curve for global optimization problems
\thanks{This study was supported by the Russian Science Foundation, project 
No.\,16-11-10150.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set 
% an abbreviated paper title here
%
\author{Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
Ilia Lebedev\orcidID{0000-0002-8736-0652}}
%
\authorrunning{K. Barkalov, I. Lebedev}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, 
Russia \email{konstantin.barkalov@itmm.unn.ru}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{Dimensionality reduction \and Global optimization \and 
Multiextremal functions.}
\end{abstract}
%
%
%
\section{Introduction}

\section{The core global search algorithm}

According to the algorithm, the first two trials are executed at the ends of 
the interval  $[a,b]$, i.e., $x^0=a,\;x^1=b$. The function values $z^0=\varphi
(x^0),\;z^1=\varphi(x^1)$  are computed and the number $k$ is set to 1. In 
order to select the point of a new trial $x^{k+1}, k\geq 1,$  it is necessary 
to perform the following steps.

\textbf{Step 1.} Renumber by subscripts (beginning from zero) the points $x^i,
\:0\leq i\leq k$, of the previous trials in increasing order, i.e.,
\begin{displaymath}
a=x_0<x_1<\ldots <x_{k}=b.
\end{displaymath} 
Juxtapose to the points $x_i,\:1\leq i\leq k$,  the function values $z_i=
\varphi(x_i),\:0\leq i\leq k$.

\textbf{Step 2.} Compute the maximum absolute value of the first divided 
differences 
\begin{displaymath}
\mu=\max_{1\leq i\leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i}
\end{displaymath}
where $\Delta_i = x_i-x_{i-1}$. If the above formula yields a zero value, 
assume that $\mu = 1$.

\textbf{Step 3.} For each interval $(x_{i-1},x_i),1\leq i\leq k$,  calculate 
the characteristic
\begin{displaymath}
R(i)=r\mu(x_i-x_{i-1})+\frac{(z_i-z_{i-1})^2}{r\mu(x_i-x_{i-1})}-2(z_i+z_{i-1}
),
\end{displaymath} 
where $r>1$ is a predefined parameter of the method.

\textbf{Step 4.} Find the interval $(x_{t-1},x_t)$ with the maximum 
characteristic
\begin{equation}\label{MaxR}
R(t)=\max_{1\leq i\leq {k}}R(i).
\end{equation}

\textbf{Step 5.} Execute the new trial at the point 
\begin{displaymath}
x^{k+1}=\frac{1}{2}(x_{t-1}+x_t) - \frac{z_t-z_{t-1}}{2r\mu}.
\end{displaymath}

The algorithm terminates if the condition $\Delta_t<\epsilon$ is satisfied; 
here $t$ is from (\ref{MaxR}) and $\epsilon>0$ is the preset accuracy. For 
estimation of the global solution, values
\[
z_k^\ast=\min_{0\leq i \leq k}\varphi(x^i), \ x_k^\ast=\arg \min_{0\leq i \leq
 k}\varphi(x^i).
\]
are selected.

\section{Dimensionality reduction}
\subsection{Dimensionality reduction using Peano-type space-filling curves}

\subsection{Nested optimization schemes}

The nested optimization scheme of dimensionality reduction is based on the 
well-known relation (see, e.g., [])
\begin{equation}\label{nested}
\min_{y \in D}\varphi(y) = \min_{a_1\leq y_1 \leq b_1}\min_{a_2\leq y_2 \leq 
b_2}...\min_{a_N\leq y_N \leq b_N}\varphi(y),
\end{equation}
which allows replacing the solving of multidimensional problem (\ref{multi_
problem}) by solving a family of one-dimensional subproblems related to each 
other recursively.

In order to describe the scheme let us introduce a set of reduced functions 
as follows:
\begin{equation}\label{nested_N}
\varphi^N(y_1,...,y_N) = \varphi(y_1,...,y_N),
\end{equation}
\begin{equation}\label{nested_i}
\varphi^i(y_1,...,y_i) = \min_{a_{i+1}\leq y_{i+1}\leq b_{i+1}} \varphi^{i+1}(
y_1,...,y_i,y_{i+1}), 1\leq i\leq N-1.
\end{equation}

Then, according to relation (\ref{nested}), solving of multidimensional 
problem (\ref{multi_problem}) is reduced to solving a one-dimensional problem 
\begin{equation}\label{nested_1}
%\varphi^1(y_1^\ast) = 
\min_{a_1\leq y_1\leq b_1}\varphi^1(y_1).
\end{equation}
But in order to evaluate the function $\varphi^1$ at a fixed point $y_1$ it 
is necessary to solve the one-dimensional problem of the second level
\begin{equation}
\varphi^1(y_1) = \min_{a_2\leq y_2\leq b_2}\varphi^2(y_1,y_2),
\end{equation}
and so on up to the univariate minimization of the function $\varphi^N(y_1
,...,y_N)$ with fixed coordinates $y_1,...,y_{N-1}$ at the $N$-th level of 
recursion.

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
 \bibliographystyle{splncs04}
 \bibliography{bibliography}

\end{document}

### 2019-01-17 2019/WCGO/WCGO.tex
All the text except the Results

@@
 
+%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
+\usepackage[utf8x]{inputenc}
+\usepackage[english,russian]{babel}
+\usepackage{cmap}
+%%%%
+
 % If you use the hyperref package, please uncomment the following line
@@
 %
-\title{ Generalized scheme for dimensionality reduction based on nested 
-optimization and Peano curve for global optimization problems
+\title{Generalized dimensionality reduction scheme 
+for global optimization problems
 \thanks{This study was supported by the Russian Science Foundation, project 
@@
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Generalized dimensionality reduction scheme}
 % If the paper title is too long for the running head, you can set 
@@
 %
-\author{Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
-Ilia Lebedev\orcidID{0000-0002-8736-0652}}
+\author{Konstantin Barkalov %\orcidID{0000-0001-5273-2471}  
+\and Ilia Lebedev %\orcidID{0000-0002-8736-0652}
+}
 %
@@
 %
-\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, 
+\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, 
 Russia \email{konstantin.barkalov@itmm.unn.ru}}
@@
 \section{Introduction}
+This paper considers ''black-box'' global optimization problems of the following form:
+\begin{eqnarray}\label{main_problem}
+& \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
+& D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
+\end{eqnarray}
+The objective function is assumed to satisfy the Lipschitz condition 
+\[
+\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in D,\; 0<L<\infty,
+\]
+with the constant $L$ unknown a priori.
 
-\section{The core global search algorithm}
+\Russian
 
+Известным способом решения многоэкстремальных задач являются мультистартовые схемы (см., например, [8–10]).
+В этих схемах в области поиска размещается некоторая сетка, точки которой использутся как стартовые поиска экстремума локальным методом, а затем выбирается наименьший из найденных экстремумов.
+Отдельной проблемой в рамках данного подхода является выбор стартовых точек, осуществляемый, как правило, на основе метода Монте-Карло \cite{Zhigljavsky2008}.
+Этот подход вполне применим для задач с небольшим числом локальных минимумов, имеющих широкие области притяжения, но для задач с существенной многоэкстремальностью его эффективность резко падает.
+
+В настоящее время для решения задач глобальной оптимизации широко используются генетические алгоритмы (см., например, \cite{Yang2013}), которые так или иначе основаны на идеях случайного поиска.
+В силу простоты реализации и использования они получили большую популярность, однако по качеству работы (численной оценкой которого может служить число корректно решенных задач из некоторого тестового набора) они существенно уступают детерминированным алгоритмам \cite{Kvasov2018,Sergeyev2018}.
+
+Если говорить о детерминированных методах глобальной оптимизации, то многие методы данного класса основаны на различных способах разбиения области поиска на систему подобластей и последующего выбора наиболее перспективной подобласти для размещения очередного испытания (вычисления целевой функции).
+Результаты в данном направлении представлены в работах \cite{Zilinskas2010,Paulavicius2016,Evtushenko2013,Jones2009,Sergeyev2015}.
+
+Наконец, для разработки методов многомерной оптимизации широко применяется подход, связанный с редукцией многомерной задачи к эквивалентной одномерной или системе одномерных подзадач и последующим решением одномерных задач эффективными методами оптимизации функций одной переменной.
+Используются две такие схемы: редукция на основе кривых, заполняющих пространство (кривых Пеано, или разверток) \cite{Strongin2000,Sergeyev2013}, и схема рекурсивной вложенной оптимизации \cite{Strongin2000,Grishagin2001}.
+В статье \cite{Grishagin2016} предложена адаптивная схема редукции, обобщающая классическую схему рекурсивной оптимизации.
+Адаптивная схема существенно повышает эффективность оптимизации по сравнению с базовым прототипом \cite{Grishagin2016_1}.
+В данной работе предлагается обобщение адаптивной схемы редукции размерности, комбинирующее использование вложенной оптимизации и кривых Пеано.
+При таком подходе вложенные подзадачи в адаптивной схеме могут быть как одномерными, так и многомерными.
+В последнем случае для редукции размерности вложенных подзадач используются развертки.
+
+\section{The global search algorithm}
+\label{SectionCore}
+
+As a core problem we consider a one-dimensional multiextremal optimization 
+problem
+\begin{equation}\label{uni_problem}
+\varphi^\ast = \varphi(x^\ast)=\min{\left\{\varphi(x):x\in \left[0,1\right] 
+\right\}}
+\end{equation}
+with objective function satisfying the Lipschitz condition.
+
+Let us give the description of the global search algorithm (GSA) applied for 
+solving the above problem (according to \cite{Strongin2000}). 
+GSA involves constructing a sequence of points $x^i$, where the values of the 
+objective function $z^i=\varphi(x^i)$ are calculated. Let us call the 
+function value calculation process the \textit{trial}. 
 According to the algorithm, the first two trials are executed at the ends of 
@@
 \:0\leq i\leq k$, of the previous trials in increasing order, i.e.,
-\begin{displaymath}
+\[
 a=x_0<x_1<\ldots <x_{k}=b.
-\end{displaymath} 
-Juxtapose to the points $x_i,\:1\leq i\leq k$,  the function values $z_i=
-\varphi(x_i),\:0\leq i\leq k$.
+\] 
+Juxtapose to the points $x_i, 0\leq i\leq k$,  the function values $z_i=
+\varphi(x_i), 0\leq i\leq k$.
 
@@
 differences 
-\begin{displaymath}
+\begin{equation}\label{mu}
 \mu=\max_{1\leq i\leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i}
-\end{displaymath}
+\end{equation}
 where $\Delta_i = x_i-x_{i-1}$. If the above formula yields a zero value, 
@@
 the characteristic
-\begin{displaymath}
-R(i)=r\mu(x_i-x_{i-1})+\frac{(z_i-z_{i-1})^2}{r\mu(x_i-x_{i-1})}-2(z_i+z_{i-1}
-),
-\end{displaymath} 
+\begin{equation}\label{R}
+R(i)=r\mu\Delta_i+\frac{(z_i-z_{i-1})^2}{r\mu\Delta_i}-2(z_i+z_{i-1}),
+\end{equation} 
 where $r>1$ is a predefined parameter of the method.
@@
 \textbf{Step 5.} Execute the new trial at the point 
-\begin{displaymath}
+\begin{equation}\label{xk1}
 x^{k+1}=\frac{1}{2}(x_{t-1}+x_t) - \frac{z_t-z_{t-1}}{2r\mu}.
-\end{displaymath}
+\end{equation}
 
@@
 here $t$ is from (\ref{MaxR}) and $\epsilon>0$ is the preset accuracy. For 
-estimation of the global solution, values
+estimation of the global solution values
 \[
@@
 are selected. 
+The theory of algorithm convergence is presented in \cite{Strongin2000}.
 
@@
 \subsection{Dimensionality reduction using Peano-type space-filling curves}
+\label{SectionPeano}
 
-\subsection{Nested optimization schemes}
+The use of Peano curve $y(x)$ 
+\begin{equation}
+\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq N\right\}=\left\{
+y(x):0\leq x \leq 1 \right\}
+\end{equation}
+unambiguously mapping the interval of real axis $[0,1]$ onto a 
+$N$-dimensional cube is the first of the dimension reduction methods considered.
+Problems of numerical construction of Peano-type space-filling curves and the 
+corresponding theory are considered in detail in \cite{Strongin2000}, 
+\cite{Sergeyev2013}. Here we will note that a numerically 
+constructed curve (\textit{evolvent}) is $2^{-m}$ accurate approximation of 
+the theoretical Peano curve, where $m$ is an evolvent construction parameter.
+
+By using this kind of mapping it is possible to reduce the multidimensional 
+problem~(\ref{main_problem}) to a univariate problem
+\begin{equation}
+\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]
+\right\}}.
+\end{equation}
+An important property of such mapping is preservation of boundedness of 
+function relative differences (see \cite{Strongin2000,Sergeyev2013}). If the 
+function $\varphi(y)$ in the domain $D$ satisfies the Lipschitz condition, 
+then the function $\varphi(y(x))$ on the interval $[0,1]$ will satisfy a 
+uniform H{\"o}lder condition
+\begin{equation}\label{Holder}
+\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N},
+\end{equation}
+where the H{\"o}lder constant $H$ is linked to the Lipschitz constant $L$ by 
+the relation
+\begin{equation}
+H=2L\sqrt{N+3}.
+\end{equation}
+Condition (\ref{Holder}) allows adopting the algorithm for solving the 
+one-dimensional problems presented in Section \ref{SectionCore} for solving the 
+multidimensional problems reduced to the one-dimensional ones. For this, the 
+lengths of intervals $\Delta_i$  involved into rules (\ref{mu}),(\ref{R}) of the
+algorithm are substituted by the lengths 
+\begin{equation}
+\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}
+\end{equation}
+and the following expression is introduced instead of formula (\ref{xk1}):
+\begin{equation}
+x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}
+\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N.
+\end{equation}
+
+\subsection{Nested optimization scheme}
 
@@
 \end{equation}
-which allows replacing the solving of multidimensional problem (\ref{multi_
-problem}) by solving a family of one-dimensional subproblems related to each 
-other recursively.
+which allows replacing the solving of multidimensional problem 
+(\ref{main_problem}) by solving a family of one-dimensional subproblems 
+related to each other recursively.
 
@@
 Then, according to relation (\ref{nested}), solving of multidimensional 
-problem (\ref{multi_problem}) is reduced to solving a one-dimensional problem 
+problem (\ref{main_problem}) is reduced to solving a one-dimensional problem 
 \begin{equation}\label{nested_1}
@@
 
+For the nested scheme presented above, a generalization (\textit{block nested 
+optimization scheme}), which combines the use of evolvents and the nested 
+scheme has been elaborated in \cite{Barkalov2016}.
+
+Let us consider vector $y$ as a vector of block variables
+\begin{equation}
+y=(y_1,y_2,...,y_N)=(u_1,u_2,...,u_M),
+\end{equation}
+where the $i$-th block variable $u_i$ is a vector of vector $y$ components, 
+taken serially i.e. $u_1=(y_1,y_2,...,y_{N_1})$, $u_2=(y_{N_1+1},y_{N_1+2}
+,...,y_{N_1+N_2})$,..., $u_M=(y_{N-N_M+1},y_{N-N_M+2},...,y_{N})$, where $N_1+
+N_2+...+N_M=N$.
+
+Using the new variables, main relation of the nested scheme (\ref{nested}) 
+can be rewritten in the form 
+\begin{equation}\label{block_nested}
+\min_{y \in D}\varphi(y) = \min_{u_1\in D_1}\min_{u_2\in D_2}...\min_{u_M \in 
+D_M}\varphi(y),
+\end{equation}
+where the subdomains $D_i, 1 \leq i \leq M$, are projections of initial 
+search domain $D$ onto the subspaces corresponding to the variables $u_i, 1 
+\leq i \leq M$.
+
+The formulae defining the method of solving of problem (\ref{main_problem}) 
+based on relation (\ref{block_nested}), in general, are the same to the ones 
+of nested scheme (\ref{nested_N})--(\ref{nested_1}). It is only necessary to 
+substitute the original variables $y_i, 1\leq i \leq N$,  by the block 
+variables $u_i, 1\leq i \leq M$.
+
+At that, the nested subproblems 
+\begin{equation}\label{block_nested_i}
+\varphi^i(u_1,...,u_i) = \min_{u^{i+1}\in D_{i+1}} \varphi_{i+1}(u_1,...,u_i,u
+_{i+1}), 1\leq i\leq M-1.
+\end{equation}
+in the block scheme are the multidimensional ones. The dimension reduction 
+method based on Peano curves can be applied for solving these ones. It is a 
+principal difference from the initial nested scheme.
+
+\subsection{Block adaptive optimization scheme}
+
+\Russian
+Решение возникающего множества подзадач вида (\ref{nested_i}) (для nested optimization scheme) или вида (\ref{block_nested_i}) (для block nested optimization scheme) может быть организовано различными способами.
+Очевидный способ (детально проработанный в \cite{Grishagin2001,Grishagin2015} для nested optimization scheme и в \cite{Barkalov2014,Barkalov2016} для block nested optimization scheme) основан на решении поздадач в соответствии с рекурсивным порядком их порождения.
+Однако здесь возникает потеря значительной части информации о objective function при решении многомерной задачи.
+Иным  подходом является адаптивная схема, в которой все подзадачи решаются одновременно, что позволяет более полно учитывать информацию о многомерной задаче и за счет этого ускорять процесс ее решения.
+
+Для случая одномерных поздадач данный подход был реализован и теоретически обоснован в \cite{Grishagin2016,Grishagin2016_1,Grishagin2018}.
+Настоящая работа предлагает обобщение адаптивной схемы для случая многомерных подзадач.
+Дадим краткое описание ее основных элементов.
+
+Пусть вложенные подзадачи вида (\ref{block_nested_i}) решаются с помощью многомерного алгоритма глобального поиска, описанного в Section \ref{SectionPeano}.
+Тогда каждой подзадаче (\ref{block_nested_i}) можно присвоить некоторое числовое
+значение, называемое характеристикой этой задачи.
+К качестве такой характеристики можно взять значение from (\ref{MaxR}), т.е.
+максимальную характеристику интервалов, сформированных в данной задаче.
+В соответствии с правилом вычисления характеристик (ссылка), чем выше значение характеристики, тем более перспективной является подзадача для продолжения поиска в ней глобального минимума исходной задачи (ссылка)  .
+Поэтому на каждой итерации выбирается подзадача с максимальной характеристикой для проведения в ней очередного испытания.
+Это испытание либо вычисляет значение целевой функции $\varphi(y)$ (если выбранная подзадача принадлежала уровню $j=M$), либо порождает новые подзадачи согласно (\ref{block_nested_i}) при $j\leq M-1$.
+В последнем случае новые порожденные задачи добавляются к текущему множеству задач, вычисляются их характеристики и процесс повторяется.
+Завершение процесса оптимизации происходит, когда в корневой задаче выполняется условие остановки алгоритма, решающего эту задачу.
+
+\section{Results of numerical experiments}
+
+One of the well-known approaches to investigating and comparing multiextremal 
+optimization algorithms is based on the application of these methods for 
+solving a set of test problems generated randomly.
+
+The numerical comparison of the algorithms has been carried out using the Grishagin test problems [] and the GKLS test problem generator \cite{Gaviano2003}.
+
+\section{Conclusion}
+
 %

### 2019-01-18 2019/WCGO/WCGO.tex
All except results and conclusion

@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-150--250 words.
 
-\keywords{Dimensionality reduction \and Global optimization \and 
-Multiextremal functions.}
+В статье рассматриваются задачи многомерной многоэкстремальной оптимизации и численные методы их решения.
+Об оптимизируемой функции делается лишь общее предположение, что она удовлетворяет условию Липшица с неизвестной априори константой.
+Задачи такого типа часто встречаются в приложениях.
+Рассмотрено два подхода к редукции размерности в задачах многомерной оптимизации.
+Первый из них использует Peano-type space-filling curves, отображающие одномерный отрезок на многомерную область.
+Второй основан на nested optimization scheme, that reduces the mul- tidimensional problem to a family of one-dimensional subproblems.
+Предложена обобщенная схема, комбинирующая эти два подхода.
+В новой схеме решение многомерной задачи сводится к решению семейства задач меньшей размерности, в которых, в свою очередь, используются space-filling curves.
+Реализован адаптивный алгоритм, в котором все возникающие подзадачи решаются одновременно.
+Проведены численные эксперименты на нескольких сотнях тестовых задач, подтверждающие эффективность предложенной обобщенной схемы.
+
+\keywords{Global optimization \and Multiextremal functions \and 
+Dimensionality reduction \and Peano curve \and Nested optimization \and 
+Numerical methods.}
 \end{abstract}
@@
 \Russian
-
-Известным способом решения многоэкстремальных задач являются мультистартовые схемы (см., например, [8–10]).
+Известным способом решения многоэкстремальных задач являются мультистартовые схемы.
 В этих схемах в области поиска размещается некоторая сетка, точки которой использутся как стартовые поиска экстремума локальным методом, а затем выбирается наименьший из найденных экстремумов.
@@
 \[
-a=x_0<x_1<\ldots <x_{k}=b.
+0=x_0<x_1<\ldots <x_{k}=1.
 \] 
@@
 The nested optimization scheme of dimensionality reduction is based on the 
-well-known relation (see, e.g., [])
+well-known relation (see, e.g., \cite{Carr1964})
 \begin{equation}\label{nested}
@@
 
-One of the well-known approaches to investigating and comparing multiextremal 
-optimization algorithms is based on the application of these methods for 
-solving a set of test problems generated randomly.
+One of the well-known approaches to investigating and comparing multiextremal optimization algorithms is based on the application of these methods for solving a set of test problems generated randomly.
 
-The numerical comparison of the algorithms has been carried out using the Grishagin test problems [] and the GKLS test problem generator \cite{Gaviano2003}.
+The numerical comparison of the algorithms has been carried out using the Grishagin test problems $F_{gr}$ (описание данных функций см., например, в \cite{Grishagin1994}, test function 4) and the GKLS generator \cite{Gaviano2003}.
+
+В работах \cite{Barkalov2015,Grishagin2018} было показано, что global search algorithm как с использованием разверток, так и в сочетании с адаптивной схемой редукции размерности превосходит многие известные алгоритмы, включая методы DIRECT и DIRECT\textit{l}.
+Поэтому в данном исследовании мы ограничимся сравнением вариантов GSA с различными схемами редукции размерности.
+
+Для сравнения эффективности работы алгоритмов используются два критерия: the average number of trials and operational characteristic.
+Операционной характеристикой алгоритма называется функция $P(k)$, определяемая как доля задач, для решения которых потребовалось не более чем $k$ испытаний.
+The global minimizer $y^\ast$ was considered to be found, if the algorithm generated a trial point $y^k$ in the vicinity of the global minimum, i.e.
+$\left|y^k-y^\ast\right| < \delta \left\|b-a\right\|$, where $\delta = 0.01$, $a$ and $b$ are the boundaries of the search domain $D$.
+
+Первая серия экспериментов была проведена на двумерных задачах классов $F_{gr}$, GKLS \textit{Simple}, GKLS \textit{Hard} (100 функций каждого класса).
+В таблице 1  представлено среднее число испытаний, выполненных GSA с использованием evolvents ($K_e$), nested optimization scheme ($K_n$ ), adaptive nested optimization scheme ($K_a$).
+
+\begin{table}
+\centering
+\caption{Average number of iterations for 2D problems.}\label{tab1}
+\begin{tabular}{lccc}
+\hline\noalign{\smallskip}
+ &  $F_{gr}$  &  GKLS \textit{Simple} &  GKLS \textit{Hard} \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+ $K_e$ & 180  & 252 & 674 \\
+ $K_n$ & 341  & 697 & 1252 \\
+ $K_a$ & 215  & 279 & 815 \\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
+Результаты экспериментов показывают, что GSA с использованием adaptive nested optimization scheme не сильно уступает по быстродействию GSA с развертками, и оба они значительно превосходят GSA в сочетании с nested optimization scheme.
 
 \section{Conclusion}
+
+В данной работе предложена обобщенная адаптивная схема редукции размерности для задач глобальной оптимизации, комбинирующая использование кривых Пеано и схему вложенной оптимизации.
+Для решения редуцированных подзадач меньшей размерности используется алгоритм глобального поиска.
+Приведена вычислительная схема алгоритма, рассмотрены алгоритмические моменты работы адаптивной схемы редукции размерности.
+Проведены вычислительные эксперименты на серии тестовых задач с целью сравнения эффективности различных схем редукции размерности. 
+Результаты экспериментов показывают, что . 
+Перспективы ...
 

### 2019-01-20 2019/MOTOR_Sovrasov/motor_paper.tex
MOTOR: add a draft of paper

--- 2018-12-03 2019/PAVT/pavt_paper.tex
@@
-%%%%%%%%%%%%%%%%%%%% author.tex
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.20 of 2017/10/04
 %
-% sample root file for your "contribution" to a proceedings volume
-%
-% Use this file as a template for your own input.
-%
-%%%%%%%%%%%%%%%% Springer
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass{svproc}
-%
-% RECOMMENDED
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
+\documentclass[runningheads]{llncs}
 %
 \usepackage{graphicx}
-\usepackage{marvosym}
 \usepackage{amssymb}
+\usepackage{subfig}
 \usepackage{cite}
-\usepackage{subfig}
-% to typeset URLs, URIs, and DOIs
-\usepackage{url}
-\usepackage{hyperref}
-\def\UrlFont{\rmfamily}
 
-\def\orcidID#1{\unskip$^{[#1]}$}
-\def\letter{$^{\textrm{(\Letter)}}$}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following line
+% to display URLs in blue roman font according to Springer's eBook style:
+% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
 \begin{document}
-\mainmatter              % start of a contribution
 %
-\title{Comparison of several sequential and parallel derivative-free global optimization algorithms}
+\title{Comparison of several derivative-free global optimization algorithms\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}}
 %
-\titlerunning{Comparison of several optimization methods}  % abbreviated title (for running head)
-%                                     also used for the TOC unless
-%                                     \toctitle is used
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
 %
-\author{Vladislav Sovrasov\letter \and Semen Bevzuk}
+\author{Vladislav Sovrasov}
 %
-\authorrunning{Vladislav Sovrasov et al.} % abbreviated author list (for running head)
-%
-%%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{Vladislav Sovrasov and Semen Bevzuk}
+\authorrunning{V. Sovrasov}
 %
 \institute{Lobachevsky State University of Nizhni Novgorod, Russia \\
-  \email{sovrasov.vlad@gmail.com, semen.bevzuk@gmail.com}
-}
-
-\maketitle              % typeset the title of the contribution
+  \email{sovrasov.vlad@gmail.com}
+}%
+\maketitle              % typeset the header of the contribution
+%
 
@@
 
-\subsection{Parallel Algorithm on Distributed Memory Exploiting a Set of Evolvents}
-\label{sub:parallel_evolvents}
-\subsubsection{Rotated Evolvents}
-One of the possible ways to overcome the negative effects of using a numerical
-approximation of evolvent (it destroys the information about the neighbor points in
-$\mathbb{R}^N$ space)
-consists in using the multiple mappings
-\begin{equation}
-  Y_L(x)=\left\{y^0(x),\ y^1(x),...,\ y^L(x)\right\}
-\end{equation}
-instead of single Peano curve $y(x)$ (see \cite{strSergGO}).
-The building of a set of Peano curves by rotation of the evolvents around the coordinate origin is a
-distinctive feature was proposed in \cite{Gergel2009}. Taking into account the initial mapping, one
-can conclude that current implementation of the
-method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain
-onto the corresponding one-dimensional intervals. This method for building a set of mappings can be
-``scaled'' easily to obtain more evolvents (up to
-$2^N$) if necessary.
-
-Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the
-problems
-\[
-\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leq s\leq S,
-\]
-on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a
-separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
-particular processor are interpreted as the results of the trials in the rest problems (in the
-corresponding points \(x^{k_1},\dots,x^{k_S})\). In this approach, a trial at the point \(x^k \in
-[0,1]\) executed in the framework of the \(s^{th}\) problem, consists in the following sequence
-of
-operations:
-\par
-1. Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
-\par
-2. Inform the rest of processors about the start of the trial execution at the point \( y^k\) (the
-blocking of the point \(y^k\) ).
-\par
-3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leq s\leq S\), of the point \(y^k\) and
-interpret the
-trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
-\(x{}^{k_1} ,\dots,x{}^{k_s} \)
-\par
-4. Inform the rest of processors about the trial results at the point \(y^k\).
-\par
-The decision rules for the mentioned parallel algorithm, in general, are the same as the rules of the
-sequential algorithm (except the method of the trial execution). Each processor has its own copy
-of the software realizing the computations of the problem functions and the decision rule of the
-index algorithm. For the organization of the interactions among the processors, the queues are
-created on each processor, where the processors store the information on the executed iterations
-in the form of the tuples: the processor number \(s\), the trial point \(x{}^{k_s}\).
-The modifications of the method taking into account multiply criteria are given in \cite{GERGEL2017,Gergel2018}.
-\par
-The mentioned parallelization scheme was proposed in \cite{Gergel2009} and implemented in the
-Globalizer system with the use of MPI technology. Main
-features of implementation consist in the following:
-\begin{itemize}
-  \item A separate MPI-process is created for each
- of \(S\) one-dimensional problems being solved, usually, one process per one processor
- employed.
- \item Each process can use $p$ threads to parallel execute $p$ trials, usually one thread per an
-accessible core.
-\end{itemize}
-
-\subsection{MIDACO Parallel Solver}
-MIDACO (Mixed Integer Distributed Ant Colony Optimization) \cite{Schlueter2012} is a solver
-which implements the evolutionary algorithm
-the ant colony optimization metaheuristic for continuous search domains. ACO was originally
-proposed to solve the TSP, but later has been successfully adopted to solve mixed integer nonlinear
-programming problems \cite{SCHLUTER2009}.
-
-The MIDACO solver utilizes reverse communication architecture \cite{Schlueter2012}.
-Reverse communication means that the call of the objective function happens outside and
-independently of the MIDACO source code.
-Relying on this feature the following distributed parallelization scheme was implemented:
-\begin{enumerate}
-  \item MIDACO runs at the master node and generates \(n \times p\) new trial points at each
-iteration.
-  \item Each of \(n\) distributed computational nodes gets \(p\) points from the master and performs
-\(p\) parallel trials on local computing devices.
-  \item Each of \(n\) distributed computational nodes sends \(p\) values of the objective function to
-the master.
-\end{enumerate}
-
-In this scheme all distributed communications were implemented using the MPI library.
-Parallelism within a single node is powered by the OpenMP standard.
-
 \subsection{Sequential Methods}
@@
 \section{Results of Numerical Experiments}
-\subsection{Sequential Algorithms}
 
@@
 
-\subsection{Parallel Algorithms}
-
-Computational experiments with parallel algorithms have been carried out on the Lobachevsky
-supercomputer at State University of Nizhni Novgorod. A computational node includes 2 Intel
-Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. Each CPU has 8 cores (i. e. total 16
-cores were available per a node). All the computational nodes are working under CentOS Linux 7.2.
-All the considered algorithms are implemented using C++ and compiled by the GCC 4.8.5.
-
-The comparison of two parallel algorithms implemented in the Globalizer and MIDACO systems
-was conducted on four-dimensional GKLS problems since the problems of larger dimensionality are
-not supported by the free version of MIDACO. As in the case of the sequential algorithms, the
-parameters of methods were determined for each problem class. For MIDACO, the parameter
-\(focus = -1\). In the case of Globalizer, the parameter \(r\) was selected following Table
-\ref{tab:params}. This parameter decreased by 0.3 with doubling the number of employed evolvents
-since the increasing of the number of evolvents enhances the reliability of the method from Section
-\ref{sub:parallel_evolvents} as well (see \cite{SOVRASOV2018} for details).
-
-If one considers the costs of parallelism to be negligible as compared to the costs of computing the
-objective functions in the optimization problems, the speedup in time due to the use of the parallel
-method would be equal to the speedup with respect to the number of iterations. However, actually
-this assumption is true if the computing of the objective function is well computation costly only. In
-all numerical experiments, the time of computing the objective function was approximately $2.5\times
-10^{-3}$ s.
-
-In Table~\ref{tab:iterations}, an averaged number of iterations when solving 100 problems from
-each considered class is presented.
-The number of iterations is reduced considerably with increasing the number of nodes and the
-number of threads on each node. Also, the number of solved problems in each class is given  in
-brackets.
-Globalizer solves almost all problems in both classes whereas MIDACO hasn't solved a considerable
-number of problems from the GKLS 4d Hard class. Also, it is worth noting that the reliability of
-MIDACO decreased with increasing number of nodes on both problem classes whereas the
-reliability of Globalizer remains stable.
-
-\begin{table}
-  \centering
-  \caption{Averaged numbers of iterations executed by the parallel algorithms for solving the test
-optimization problems}
-  \label{tab:iterations}
-  \begin{tabular}{cccccccc}
-    \cline{3-8}\noalign{\smallskip}
-    \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{Globalizer} & &
-\multicolumn{2}{c}{MIDACO}   \\
-    \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
-    \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
-\textit{Hard}  \\
-    \noalign{\smallskip}\hline
-    I & \textbf{1 cluster node}
-      & \textit{1} &   25270 (100)  & 55180 (99) & & 27645 (98) & 72068 (71)  \\
-    &  & \textit{16} & 1765 (100)  & 3714 (100) & &  1640 (97) & 4304 (70) \\
-    \hline \noalign{\smallskip}
-II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
-  & \textit{1} & 13056 (100) & 22938 (99) & & 10558 (89) & 27128 (73) \\
-&   & \textit{16} & 732 (100) & 1759 (100)  &  & 1130 (92) & 2254 (73) \\
-    \hline \noalign{\smallskip}
-III & \textbf{4 cluster nodes} %\multirow{3}{*}{}
-  & \textit{1}  & 5016 (100) & 12703 (100) & & 5777 (94) & 15980 (75) \\
-& & \textit{16} & 367 (100) & 776 (100) & & 529 (88) &  1264 (66) \\
-    \hline \noalign{\smallskip}
-VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-  & \textit{1}  & 2103 (100) & 5063 (100) & & 2847 (97) & 9853 (83)\\
-& & \textit{16} & 145 (100)  & 310 (100)  & & 272 (83) & 774 (57)\\
-    \hline \noalign{\smallskip}
-V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
-  & \textit{1}  & 1155 (100) & 2399 (100) & & 2233 (98) & 6022 (86) \\
-& & \textit{16} & 76 (100)  & 159 (100)  & & 168 (87) & 393 (53)\\
-    \noalign{\smallskip}\hline
-  \end{tabular}
-\end{table}
-
-The averaged speedup in the number of iterations and the one in time (in brackets) are presented in
-Table \ref{tab:speedup}. In the first row of Table \ref{tab:speedup}, the averaged number of
-iterations and the averaged time of solving the problems in the sequential mode (in brackets) are
-presented. MIDACO uses a smaller number of communications in the parallel computations than the Globalizer.
-Therefore, the indicators of speedup increased monotonously with increasing number of
-computational nodes employed. In Globalizer, each node should transfer the results of the executed
-iterations to all other nodes. Therefore, the efficiency of parallelization decreased with increasing
-number of nodes and the number of active threads on each node. This is evident from Table
-\ref{tab:speedup} and Picture \ref{fig:speedup_p16}: the speedup in time decreased at the transition from 4 nodes to 8 ones when
-using 16 threads. In the case of using 1 thread per a node, the speedups of both optimization methods
-were approximately the same (see Picture \ref{fig:speedup_p16}). However, it should be noted that Globalizer executed less number of
-trials, and spends much less time for solving all problems when \(p=1\). With increasing number of
-nodes and threads, MIDACO approaches Globalizer in the averaged time of problem solving (see Pictures \ref{fig:time_p1}, \ref{fig:time_p16}),
-however, its reliability decreases at the same time.
-
-An anomalous superlinear speedup of MIDACO in time on the GKLS 2d Hard class when
-employing 8 and 12 nodes can be explained by the fact that, according to Table \ref{tab:iterations},
-MIDACO has solved a greater number of problems in the parallel mode than in the sequential one,
-hence, has worked until the limit of trials is expired on a less number of problems.
-
-\begin{figure}[ht]
-  \centering
-  \subfloat[$p=1$]{{\includegraphics[width=.5\textwidth]{images/speedup_midaco_globalizer_p_1.pdf}}\label{fig:speedup_p1}}
-  \subfloat[$p=16$]{{\includegraphics[width=.5\textwidth]{images/speedup_midaco_globalizer_p_16.pdf}}\label{fig:speedup_p16}}
-  \caption{Speedup demonstrated by the parallel algorithms when solving problems from the GKLS 4d Hard class}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \subfloat[$p=1$]{{\includegraphics[width=.5\textwidth]{images/time_midaco_globalizer_p_1.pdf}}\label{fig:time_p1}}
-  \subfloat[$p=16$]{{\includegraphics[width=.5\textwidth]{images/time_midaco_globalizer_p_16.pdf}}\label{fig:time_p16}}
-  \caption{Averaged execution time of the parallel algorithms when solving problems from the GKLS 4d Hard class}
-\end{figure}
-
-\begin{table}
-  \centering
-  \caption{Speedup of parallel computations executed by the parallel algorithms}
-  \label{tab:speedup}
-  \begin{tabular}{cccccccc}
-    \cline{3-8}\noalign{\smallskip}
-    \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{Globalizer} & &
-\multicolumn{2}{c}{MIDACO}   \\
-    \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
-    \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
-\textit{Hard}  \\
-    \noalign{\smallskip}\hline
-I  & \textbf{1 cluster node}  %\multirow{3}{*}{}
-    & \textit{1}   & 25270 (27.3s) & 55180 (61.8s) & & 27645 (32.2s) & 72068 (132.5s)  \\
-  &  & \textit{16} & 14.3(11.7) & 14.9(12.6)  & &  16.9(14.4) & 16.7(14.4) \\
-  \hline \noalign{\smallskip}
-II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
-  & \textit{1}      &   1.9(1.9) & 2.4(2.2)  & & 2.6(1.7) & 2.7(2.3) \\
-  &   & \textit{16} & 34.5(20.4) & 31.4(18.8) & & 24.5(18.8) & 32.0(29.1) \\
-    \noalign{\smallskip}\hline	\noalign{\smallskip}
-III  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
-& \textit{1}      & 5.0(4.6) & 4.3(4.1) & & 4.8(3.9) & 4.5(4.4) \\
-&   & \textit{16} & 68.8(23.7) & 71.2(25.8) & & 52.3(32.9) & 57.0(50.2) \\
-  \noalign{\smallskip}\hline	\noalign{\smallskip}
-VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-& \textit{1}    & 12.0(8.7) & 10.9(8.3) & & 9.7(8.2)    & 7.3(9.0)    \\
-& & \textit{16} & 174.1(4.7) & 177.8(5.3) & & 101.6(51.6) & 93.1(84.1) \\
-    \noalign{\smallskip}\hline
-    V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
-    & \textit{1}    & 21.9(11.4)  & 23.0(12.8)  & & 12.4(11.1)  & 12.0(14.8)  \\
-    & & \textit{16} & 333.5(2.7) & 347.9(3.0) & & 164.5(82.0) & 183.2(129.7) \\
-        \noalign{\smallskip}\hline
-  \end{tabular}
-\end{table}
-
 \section{Conclusions}
@@
 
-\section*{Acknowledgements}
-The study was supported by the Russian Science Foundation, project No 16-11-10150.
-
 % ---- Bibliography ----
 %
-\bibliographystyle{spmpsci}
+\bibliographystyle{splncs04}
 \bibliography{bibliography}{}
+%
 

### 2019-01-21 2019/WCGO/WCGO.tex
New figures

@@
 
-Для сравнения эффективности работы алгоритмов используются два критерия: the average number of trials and operational characteristic.
-Операционной характеристикой алгоритма называется функция $P(k)$, определяемая как доля задач, для решения которых потребовалось не более чем $k$ испытаний.
+Для сравнения эффективности работы алгоритмов используются два критерия: the average number of trials and operating characteristic.
+Операционной характеристикой алгоритма называется функция $P(k)$, определяемая как доля задач из рассматриваемой серии, для решения которых потребовалось не более чем $k$ испытаний.
 The global minimizer $y^\ast$ was considered to be found, if the algorithm generated a trial point $y^k$ in the vicinity of the global minimum, i.e.
@@
 Первая серия экспериментов была проведена на двумерных задачах классов $F_{gr}$, GKLS \textit{Simple}, GKLS \textit{Hard} (100 функций каждого класса).
-В таблице 1  представлено среднее число испытаний, выполненных GSA с использованием evolvents ($K_e$), nested optimization scheme ($K_n$ ), adaptive nested optimization scheme ($K_a$).
+В Table~\ref{tab1}  представлено среднее число испытаний, выполненных GSA с использованием evolvents ($K_e$), nested optimization scheme ($K_n$ ), adaptive nested optimization scheme ($K_a$).
+На Fig.~\ref{fig1}, Fig.~\ref{fig2}(a,b) приведены operating characteristics алгоритмов, полученные на классах $F_{gr}$, GKLS \textit{Simple}, GKLS \textit{Hard} соответственно.
+Непрерывная линия соответсвует алгоритму с использованием evolvents, короткий пунктир -- adaptive nested optimization scheme, длинный пунктир -- nested optimization scheme.
+Результаты экспериментов показывают, что GSA с использованием adaptive nested optimization scheme показывает одинаковое быстродействие  по сравнению с GSA с развертками, и оба они значительно превосходят алгоритм, использующий nested optimization scheme.
+Поэтому в дальнейших экспериментах данная схема использоваться не будет.
 
@@
 
-Результаты экспериментов показывают, что GSA с использованием adaptive nested optimization scheme не сильно уступает по быстродействию GSA с развертками, и оба они значительно превосходят GSA в сочетании с nested optimization scheme.
+\begin{figure}
+\centering
+\includegraphics[width=0.50\textwidth]{2D.pdf}
+\caption{Operating characteristics using $F_{gr}$ class} 
+\label{fig1}
+\end{figure}
+
+\begin{figure}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{2DSimple.pdf} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{2DHard.pdf} \\ (b)}
+\end{minipage}
+\caption{Operating characteristics using 2d GKLS \textit{Simple} (a) and GKLS \textit{Hard} (b) classes}
+\label{fig2}
+\end{figure}
+
+Вторая серия экспериментов была проведена на четырехмерных задачах классов GKLS \textit{Simple}, GKLS \textit{Hard} (100 функций каждого класса).
+В Table~\ref{tab2}  представлено среднее число испытаний, выполненных GSA с использованием evolvents ($K_e$), adaptive nested optimization scheme ($K_a$ ), block adaptive nested optimization scheme ($K_ba$).
+На Fig.~\ref{fig3}(a,b) приведены operational characteristics алгоритмов, полученные на классах GKLS \textit{Simple}, GKLS \textit{Hard} соответственно.
+Непрерывная линия соответсвует алгоритму с использованием evolvents, короткий пунктир -- adaptive nested optimization scheme, длинный пунктир -- block adaptive nested optimization scheme.
+Результаты экспериментов показывают,
+
+\begin{table}
+\centering
+\caption{Average number of iterations for 4D problems.}\label{tab2}
+\begin{tabular}{lcc}
+\hline\noalign{\smallskip}
+ &    GKLS \textit{Simple} &  GKLS \textit{Hard} \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+ $K_e$  &        &       \\
+ $K_a$  &  21747 & 35633 \\
+ $K_ba$ &  15942 & 33206 \\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
+\begin{figure}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{2DSimple.pdf} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{2DHard.pdf} \\ (b)}
+\end{minipage}
+\caption{Operating characteristics using 4d GKLS \textit{Simple} (a) and GKLS \textit{Hard} (b) classes}
+\label{fig3}
+\end{figure}
 

### 2019-01-22 2019/WCGO/WCGO.tex
Before translation

@@
 %
-\title{Generalized dimensionality reduction scheme 
-for global optimization problems
+\title{Generalized dimensionality reduction scheme for global optimization 
+algorithms
 \thanks{This study was supported by the Russian Science Foundation, project 
@@
 \author{Konstantin Barkalov %\orcidID{0000-0001-5273-2471}  
-\and Ilia Lebedev %\orcidID{0000-0002-8736-0652}
+\and Ilya Lebedev %\orcidID{0000-0002-8736-0652}
 }
@@
 \begin{abstract}
-
+\Russian
 В статье рассматриваются задачи многомерной многоэкстремальной оптимизации и численные методы их решения.
-Об оптимизируемой функции делается лишь общее предположение, что она удовлетворяет условию Липшица с неизвестной априори константой.
+Об оптимизируемой функции делается лишь общее предположение, что она удовлетворяет Lipschitz condition с неизвестной априори константой.
 Задачи такого типа часто встречаются в приложениях.
@@
 \section{Introduction}
-This paper considers ''black-box'' global optimization problems of the following form:
+This paper considers ``black-box'' global optimization problems of the 
+following form:
 \begin{eqnarray}\label{main_problem}
@@
 \[
-\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in D,\; 0<L<\infty,
+\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in
+ D,\; 0<L<\infty,
 \]
@@
 \begin{equation}\label{nested_1}
-%\varphi^1(y_1^\ast) = 
-\min_{a_1\leq y_1\leq b_1}\varphi^1(y_1).
+\varphi^* = \min_{a_1\leq y_1\leq b_1}\varphi^1(y_1).
 \end{equation}
@@
 \subsection{Block adaptive optimization scheme}
-
 \Russian
@@
 
-Для случая одномерных поздадач данный подход был реализован и теоретически обоснован в \cite{Grishagin2016,Grishagin2016_1,Grishagin2018}.
+Для случая одномерных поздадач данный подход был теоретически обоснован и апробирован в \cite{Grishagin2016,Grishagin2016_1,Grishagin2018}.
 Настоящая работа предлагает обобщение адаптивной схемы для случая многомерных подзадач.
@@
 значение, называемое характеристикой этой задачи.
-К качестве такой характеристики можно взять значение from (\ref{MaxR}), т.е.
+К качестве такой характеристики можно взять значение $R(t)$ from (\ref{MaxR}), т.е.
 максимальную характеристику интервалов, сформированных в данной задаче.
-В соответствии с правилом вычисления характеристик (ссылка), чем выше значение характеристики, тем более перспективной является подзадача для продолжения поиска в ней глобального минимума исходной задачи (ссылка)  .
+В соответствии с правилом вычисления характеристик (\ref{R}), чем выше значение характеристики, тем более перспективной является подзадача для продолжения поиска в ней глобального минимума исходной задачи (\ref{main_problem}).
 Поэтому на каждой итерации выбирается подзадача с максимальной характеристикой для проведения в ней очередного испытания.
@@
 
-One of the well-known approaches to investigating and comparing multiextremal optimization algorithms is based on the application of these methods for solving a set of test problems generated randomly.
+One of the well-known approaches to investigating and comparing multiextremal 
+optimization algorithms is based on the application of these methods for 
+solving a set of test problems generated randomly.
+\Russian
+The comparison of the algorithms has been carried out using the Grishagin test problems $F_{gr}$ (описание данных функций см., например, в \cite{Grishagin1994}, test function 4) and the GKLS generator \cite{Gaviano2003}.
 
-The numerical comparison of the algorithms has been carried out using the Grishagin test problems $F_{gr}$ (описание данных функций см., например, в \cite{Grishagin1994}, test function 4) and the GKLS generator \cite{Gaviano2003}.
-
-В работах \cite{Barkalov2015,Grishagin2018} было показано, что global search algorithm как с использованием разверток, так и в сочетании с адаптивной схемой редукции размерности превосходит многие известные алгоритмы, включая методы DIRECT и DIRECT\textit{l}.
+В работах \cite{Barkalov2015,Grishagin2018} было показано, что global search algorithm как с использованием разверток, так и в сочетании с адаптивной схемой редукции размерности превосходит многие известные алгоритмы глобальной оптимизации, включая методы DIRECT и DIRECT\textit{l}.
 Поэтому в данном исследовании мы ограничимся сравнением вариантов GSA с различными схемами редукции размерности.
@@
 Первая серия экспериментов была проведена на двумерных задачах классов $F_{gr}$, GKLS \textit{Simple}, GKLS \textit{Hard} (100 функций каждого класса).
-В Table~\ref{tab1}  представлено среднее число испытаний, выполненных GSA с использованием evolvents ($K_e$), nested optimization scheme ($K_n$ ), adaptive nested optimization scheme ($K_a$).
+В Table~\ref{tab1}  представлено среднее число испытаний, выполненных GSA с использованием evolvents ($K_e$), nested optimization scheme ($K_n$), adaptive nested optimization scheme ($K_a$).
 На Fig.~\ref{fig1}, Fig.~\ref{fig2}(a,b) приведены operating characteristics алгоритмов, полученные на классах $F_{gr}$, GKLS \textit{Simple}, GKLS \textit{Hard} соответственно.
 Непрерывная линия соответсвует алгоритму с использованием evolvents, короткий пунктир -- adaptive nested optimization scheme, длинный пунктир -- nested optimization scheme.
-Результаты экспериментов показывают, что GSA с использованием adaptive nested optimization scheme показывает одинаковое быстродействие  по сравнению с GSA с развертками, и оба они значительно превосходят алгоритм, использующий nested optimization scheme.
-Поэтому в дальнейших экспериментах данная схема использоваться не будет.
+Результаты экспериментов показывают, что GSA с использованием adaptive nested optimization scheme показывает примерно одинаковое быстродействие  по сравнению с GSA с развертками, и оба они значительно превосходят алгоритм, использующий nested optimization scheme.
+Поэтому в дальнейших экспериментахмы ограничимся сравнением различных вариантов адаптивной схемы редукции размерности.
 
@@
 \end{minipage}
-\caption{Operating characteristics using 2d GKLS \textit{Simple} (a) and GKLS \textit{Hard} (b) classes}
+\caption{Operating characteristics using 2d GKLS \textit{Simple} (a) and \textit{Hard} (b) classes}
 \label{fig2}
@@
 
-Вторая серия экспериментов была проведена на четырехмерных задачах классов GKLS \textit{Simple}, GKLS \textit{Hard} (100 функций каждого класса).
-В Table~\ref{tab2}  представлено среднее число испытаний, выполненных GSA с использованием evolvents ($K_e$), adaptive nested optimization scheme ($K_a$ ), block adaptive nested optimization scheme ($K_ba$).
-На Fig.~\ref{fig3}(a,b) приведены operational characteristics алгоритмов, полученные на классах GKLS \textit{Simple}, GKLS \textit{Hard} соответственно.
-Непрерывная линия соответсвует алгоритму с использованием evolvents, короткий пунктир -- adaptive nested optimization scheme, длинный пунктир -- block adaptive nested optimization scheme.
-Результаты экспериментов показывают,
+Вторая серия экспериментов была проведена на четырехмерных задачах классов GKLS \textit{Simple} and GKLS \textit{Hard} (100 функций каждого класса).
+В Table~\ref{tab2}  представлено среднее число испытаний, выполненных GSA с использованием adaptive nested optimization scheme ($K_a$), block adaptive nested optimization scheme ($K_ba$) с формированием двух уровней подзадач одинаковой размерности $N_1=N_2=2$.
+Отметим, что при использовании исходного варианта адаптивной схемы при решении задачи размерности $N=4$ формируется четыре уровня одномерных подзадач, что усложняет их обработку.
+
+На Fig.~\ref{fig3}(a,b) приведены operating characteristics алгоритмов, полученные на классах GKLS \textit{Simple} and GKLS \textit{Hard} соответственно.
+Пунктирная линия соответсвует алгоритму с использованием adaptive nested optimization scheme, непрерывная --  block adaptive nested optimization scheme.
+Результаты экспериментов показывают, что использование block adaptive nested optimization scheme дает ощутимый выигрыш по числу trials (до 35\%) по сравнению с исходной adaptive nested optimization scheme.
 
@@
 \noalign{\smallskip}\hline\noalign{\smallskip}
- $K_e$  &        &       \\
  $K_a$  &  21747 & 35633 \\
- $K_ba$ &  15942 & 33206 \\
+ $K_{ba}$ &  15942 & 33206 \\
 \noalign{\smallskip}\hline
@@
 \begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{2DSimple.pdf} \\ (a)}
+\center{\includegraphics[width=1.0\linewidth]{4DSimple.pdf} \\ (a)}
 \end{minipage}
 \begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{2DHard.pdf} \\ (b)}
+\center{\includegraphics[width=1.0\linewidth]{4DHard.pdf} \\ (b)}
 \end{minipage}
-\caption{Operating characteristics using 4d GKLS \textit{Simple} (a) and GKLS \textit{Hard} (b) classes}
+\caption{Operating characteristics using 4d GKLS \textit{Simple} (a) and \textit{Hard} (b) classes}
 \label{fig3}
@@
 \section{Conclusion}
-
-В данной работе предложена обобщенная адаптивная схема редукции размерности для задач глобальной оптимизации, комбинирующая использование кривых Пеано и схему вложенной оптимизации.
+\Russian
+В данной работе предложена обобщенная адаптивная схема редукции размерности для задач глобальной оптимизации, комбинирующая использование кривых Пеано и схему вложенной (рекурсивной) оптимизации.
 Для решения редуцированных подзадач меньшей размерности используется алгоритм глобального поиска.
-Приведена вычислительная схема алгоритма, рассмотрены алгоритмические моменты работы адаптивной схемы редукции размерности.
+Приведена вычислительная схема алгоритма, рассмотрены основные вопросы, связанные с использованием адаптивной схемы редукции размерности.
 Проведены вычислительные эксперименты на серии тестовых задач с целью сравнения эффективности различных схем редукции размерности. 
-Результаты экспериментов показывают, что . 
-Перспективы ...
+Результаты экспериментов показывают, что использование block adaptive nested optimization scheme может значительно сократить число испытаний, необходимое для решения задачи с заданной точностью. 
+Дальнейшие работы в развитии методов решения задач глобальной оптимизации, основанных на данном способе редукции размерности, могут быть связаны с использованием локальных оценок константы Липшица в различных подзадачах; схемы локального оценивания подробно рассмотрены, например, в \cite{Grishagin2015,Sergeyev2016}.
 

### 2019-01-23 2019/PAVT/pavt_paper.tex
PAVT: fix style issues

@@
 \usepackage{hyperref}
+\usepackage{caption}
+\captionsetup[table]{skip=9pt}
 \def\UrlFont{\rmfamily}
@@
 %
-\title{Comparison of several sequential and parallel derivative-free global optimization algorithms}
+\title{ComparisonOfSeveralSequentialAndParallelDerivative-freeGlobalOptimizationAlgorithms}
 %
@@
 objective function within the whole search domain
-\cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}. Nowadays, various stochastic global
+\cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}. Nowadays, various stochastic global
 optimization algorithms have attracted much attention, first of all, the evolution ones
@@
 Earlier, the comparison of the stochastic global optimization algorithms \cite{Ali2005, JSSv060i06}
-as well as of the deterministic ones \cite{posik2012, KVASOV2018245, Liberti2005} between each
+as well as of the deterministic ones \cite{posik2012, Kvasov2018, Liberti2005} between each
 other has been considered in the literature. In these works, most of modern methods have
@@
 onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
-reduction \cite{sergeyevStronginLera2013, strongin1978, strSergGO}.
+reduction \cite{Sergeyev2013, strongin1978, Strongin2000}.
 \par
@@
 The algorithms for the numerical construction of the Peano curve approximations are
-given in \cite{strSergGO}.
+given in \cite{Strongin2000}.
 
@@
 (\ref{eq:oneDimTask}) are based on the MAGS method, which can be presented as follows ---
-see \cite{strongin1978}, \cite{strSergGO}.
+see \cite{strongin1978}, \cite{Strongin2000}.
 \par
@@
 The convergence conditions and exact formulas for decision rules $R(i)$ and $d(x)$ of the
-described algorithm are given, for example, in \cite{strSergGO}.
+described algorithm are given, for example, in \cite{Strongin2000}.
 
@@
 \end{equation}
-instead of single Peano curve $y(x)$ (see \cite{strSergGO}).
+instead of single Peano curve $y(x)$ (see \cite{Strongin2000}).
 The building of a set of Peano curves by rotation of the evolvents around the coordinate origin is a

### 2019-01-25 2019/WCGO/WCGO.tex
After translation

--- 2019-01-18 2019/WCGO/WCGO.tex
@@
 %%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
-\usepackage[utf8x]{inputenc}
-\usepackage[english,russian]{babel}
-\usepackage{cmap}
+%\usepackage[utf8x]{inputenc}
+%\usepackage[english,russian]{babel}
+%\usepackage{cmap}
 %%%%
@@
 
-В статье рассматриваются задачи многомерной многоэкстремальной оптимизации и численные методы их решения.
-Об оптимизируемой функции делается лишь общее предположение, что она удовлетворяет условию Липшица с неизвестной априори константой.
-Задачи такого типа часто встречаются в приложениях.
-Рассмотрено два подхода к редукции размерности в задачах многомерной оптимизации.
-Первый из них использует Peano-type space-filling curves, отображающие одномерный отрезок на многомерную область.
-Второй основан на nested optimization scheme, that reduces the mul- tidimensional problem to a family of one-dimensional subproblems.
-Предложена обобщенная схема, комбинирующая эти два подхода.
-В новой схеме решение многомерной задачи сводится к решению семейства задач меньшей размерности, в которых, в свою очередь, используются space-filling curves.
-Реализован адаптивный алгоритм, в котором все возникающие подзадачи решаются одновременно.
-Проведены численные эксперименты на нескольких сотнях тестовых задач, подтверждающие эффективность предложенной обобщенной схемы.
+In the present paper, the multidimensional multiextremal optimization 
+problems and the numerical methods for solving these ones are considered. A 
+general assumption only is made on the objective function that this one 
+satisfies the Lipschitz condition with the Lipschitz constant not known 
+a priori. The problems of this type are frequent in the applications. 
+Two approaches to the dimensionality reduction for the multidimensional 
+optimization problems were considered. The first one uses the Peano-type 
+space-filling curves mapping a one-dimensional interval onto a multidimensional 
+domain. The second one is based on the nested optimization scheme, which 
+reduces a multi-dimensional problem to a family of the one-dimensional 
+subproblems. A generalized scheme combining these two approaches has been 
+proposed. In this novel scheme, solving a multidimensional problem is 
+reduced to solving a family of problems of lower dimensionality, in 
+which the space-filling curves are used. An adaptive algorithm, in which all 
+arising subproblems are solved simultaneously has been implemented. The 
+numerical experiments on several hundred test problems have been carried out 
+confirming the efficiency of the proposed generalized scheme.
 
@@
 
-\Russian
-Известным способом решения многоэкстремальных задач являются мультистартовые схемы.
-В этих схемах в области поиска размещается некоторая сетка, точки которой использутся как стартовые поиска экстремума локальным методом, а затем выбирается наименьший из найденных экстремумов.
-Отдельной проблемой в рамках данного подхода является выбор стартовых точек, осуществляемый, как правило, на основе метода Монте-Карло \cite{Zhigljavsky2008}.
-Этот подход вполне применим для задач с небольшим числом локальных минимумов, имеющих широкие области притяжения, но для задач с существенной многоэкстремальностью его эффективность резко падает.
+The multistart scheme is a well known method for solving the multiextremal problems.
+In such schemes, a grid is seeded in the search domain, the points of which are used as the starting ones for the search of 
+the extrema by some local method, and then the lowest of the found extrema is chosen. 
+The choice of the starting points performed, as a rule, on the basis of the Monte Carlo method is a special problem within 
+this approach \cite{Zhigljavsky2008}. 
+The approach is well applicable for the problems with a small number of local minima having a vide field of application, 
+but for the problems with essential multiextremality its efficiency falls drastically.
 
-В настоящее время для решения задач глобальной оптимизации широко используются генетические алгоритмы (см., например, \cite{Yang2013}), которые так или иначе основаны на идеях случайного поиска.
-В силу простоты реализации и использования они получили большую популярность, однако по качеству работы (численной оценкой которого может служить число корректно решенных задач из некоторого тестового набора) они существенно уступают детерминированным алгоритмам \cite{Kvasov2018,Sergeyev2018}.
+At present, the genetic algorithms one way or another based on the random search concept are used widely for solving the 
+global optimization problems (see, for example, \cite{Yang2013}). 
+Because of simplicity of implementation and usage these ones have gained a large popularity. 
+However the quality of work (the number of problems from some set solved correctly can serve as a quantitative measure 
+of which) is essentially less as compared to the deterministic algorithms \cite{Kvasov2018,Sergeyev2018}.
 
-Если говорить о детерминированных методах глобальной оптимизации, то многие методы данного класса основаны на различных способах разбиения области поиска на систему подобластей и последующего выбора наиболее перспективной подобласти для размещения очередного испытания (вычисления целевой функции).
-Результаты в данном направлении представлены в работах \cite{Zilinskas2010,Paulavicius2016,Evtushenko2013,Jones2009,Sergeyev2015}.
+If one speaks about the deterministic global optimization methods, many ones of this class are based on various methods 
+of division of the search domain into a system of subdomains followed by the selection of the most promising subdomain 
+for placing the next trial (computing the objective function). The results in this direction are presented in 
+\cite{Evtushenko2013,Jones2009,Paulavicius2016,Zilinskas2010,Sergeyev2015}.
 
-Наконец, для разработки методов многомерной оптимизации широко применяется подход, связанный с редукцией многомерной задачи к эквивалентной одномерной или системе одномерных подзадач и последующим решением одномерных задач эффективными методами оптимизации функций одной переменной.
-Используются две такие схемы: редукция на основе кривых, заполняющих пространство (кривых Пеано, или разверток) \cite{Strongin2000,Sergeyev2013}, и схема рекурсивной вложенной оптимизации \cite{Strongin2000,Grishagin2001}.
-В статье \cite{Grishagin2016} предложена адаптивная схема редукции, обобщающая классическую схему рекурсивной оптимизации.
-Адаптивная схема существенно повышает эффективность оптимизации по сравнению с базовым прототипом \cite{Grishagin2016_1}.
-В данной работе предлагается обобщение адаптивной схемы редукции размерности, комбинирующее использование вложенной оптимизации и кривых Пеано.
-При таком подходе вложенные подзадачи в адаптивной схеме могут быть как одномерными, так и многомерными.
-В последнем случае для редукции размерности вложенных подзадач используются развертки.
+Finally, the approach related to the reduction of the multidimensional problems to the equivalent one-dimensional ones (or 
+to a system of the one-dimensional subproblems) with subsequent solving the one-dimensional problems by the efficient 
+univariate optimization algorithms is widely used for the development of the multidimensional 
+optimization methods. Two such schemes are used: the reduction based on the Peano-type space-filling curves (evolvents) 
+\cite{Sergeyev2013,Strongin2000}, and the nested optimization scheme \cite{Grishagin2001,Strongin2000}.
+
+An adaptive reduction scheme generalizing the classical nested optimization one has been proposed in 
+\cite{Grishagin2016}. The adaptive scheme essentially enhances the optimization efficiency as compared to the base 
+prototype \cite{Grishagin2016_1}. In the present work, a generalization of the adaptive dimensionality reduction scheme 
+combining the use of the nested optimization and the Peano curves has been proposed. In this approach the nested 
+subproblems in the adaptive scheme can be the one-dimensional as well as the multidimensional ones. In the latter case, 
+the evolvents are used for the reduction of the dimensionality of the nested subproblems.
 
@@
 According to the algorithm, the first two trials are executed at the ends of 
-the interval  $[a,b]$, i.e., $x^0=a,\;x^1=b$. The function values $z^0=\varphi
+the interval  $[0,1]$, i.e. $x^0=0,\;x^1=1$. The function values $z^0=\varphi
 (x^0),\;z^1=\varphi(x^1)$  are computed and the number $k$ is set to 1. In 
@@
 Problems of numerical construction of Peano-type space-filling curves and the 
-corresponding theory are considered in detail in \cite{Strongin2000}, 
-\cite{Sergeyev2013}. Here we will note that a numerically 
+corresponding theory are considered in detail in 
+\cite{Sergeyev2013,Strongin2000}. Here we will note that a numerically 
 constructed curve (\textit{evolvent}) is $2^{-m}$ accurate approximation of 
@@
 An important property of such mapping is preservation of boundedness of 
-function relative differences (see \cite{Strongin2000,Sergeyev2013}). If the 
+function relative differences (see \cite{Sergeyev2013,Strongin2000}). If the 
 function $\varphi(y)$ in the domain $D$ satisfies the Lipschitz condition, 
@@
 where the $i$-th block variable $u_i$ is a vector of vector $y$ components, 
-taken serially i.e. $u_1=(y_1,y_2,...,y_{N_1})$, $u_2=(y_{N_1+1},y_{N_1+2}
+taken serially, i.e., $u_1=(y_1,y_2,...,y_{N_1})$, $u_2=(y_{N_1+1},y_{N_1+2}
 ,...,y_{N_1+N_2})$,..., $u_M=(y_{N-N_M+1},y_{N-N_M+2},...,y_{N})$, where $N_1+
@@
 
-The formulae defining the method of solving of problem (\ref{main_problem}) 
+The formulae defining the method for solving the problem (\ref{main_problem}) 
 based on relation (\ref{block_nested}), in general, are the same to the ones 
@@
 \varphi^i(u_1,...,u_i) = \min_{u^{i+1}\in D_{i+1}} \varphi_{i+1}(u_1,...,u_i,u
-_{i+1}), 1\leq i\leq M-1.
+_{i+1}), 1\leq i\leq M-1,
 \end{equation}
@@
 
-\Russian
-Решение возникающего множества подзадач вида (\ref{nested_i}) (для nested optimization scheme) или вида (\ref{block_nested_i}) (для block nested optimization scheme) может быть организовано различными способами.
-Очевидный способ (детально проработанный в \cite{Grishagin2001,Grishagin2015} для nested optimization scheme и в \cite{Barkalov2014,Barkalov2016} для block nested optimization scheme) основан на решении поздадач в соответствии с рекурсивным порядком их порождения.
-Однако здесь возникает потеря значительной части информации о objective function при решении многомерной задачи.
-Иным  подходом является адаптивная схема, в которой все подзадачи решаются одновременно, что позволяет более полно учитывать информацию о многомерной задаче и за счет этого ускорять процесс ее решения.
+The solving of the arising set of subproblems (\ref{nested_i}) (for the nested optimization scheme) or 
+(\ref{block_nested_i}) (for the block nested optimization scheme) can be organized in various ways. 
+A straightforward way (developed in details for the nested optimization scheme \cite{Grishagin2015,Grishagin2001} 
+and for the block nested optimization scheme \cite{Barkalov2016,Barkalov2014}) is based on solving the 
+subproblems according to the generation order. However, here a loss of a considerable part of the information 
+on the objective function takes place when solving the multidimensional problem. Another approach is the adaptive 
+scheme, in which all subproblems are solved simultaneously, that allows more complete accounting for the information on 
+the multidimensional and accelerating the process of its solving.
 
-Для случая одномерных поздадач данный подход был реализован и теоретически обоснован в \cite{Grishagin2016,Grishagin2016_1,Grishagin2018}.
-Настоящая работа предлагает обобщение адаптивной схемы для случая многомерных подзадач.
-Дадим краткое описание ее основных элементов.
+For the case of the one-dimensional subproblems the adaptive scheme was theoretically substantiated and tested in 
+\cite{Grishagin2016,Grishagin2016_1,Grishagin2018}. The present work proposes a generalization of the adaptive 
+scheme for the case of the multidimensional subproblems. Let us give a brief description of its basic elements.
 
-Пусть вложенные подзадачи вида (\ref{block_nested_i}) решаются с помощью многомерного алгоритма глобального поиска, описанного в Section \ref{SectionPeano}.
-Тогда каждой подзадаче (\ref{block_nested_i}) можно присвоить некоторое числовое
-значение, называемое характеристикой этой задачи.
-К качестве такой характеристики можно взять значение from (\ref{MaxR}), т.е.
-максимальную характеристику интервалов, сформированных в данной задаче.
-В соответствии с правилом вычисления характеристик (ссылка), чем выше значение характеристики, тем более перспективной является подзадача для продолжения поиска в ней глобального минимума исходной задачи (ссылка)  .
-Поэтому на каждой итерации выбирается подзадача с максимальной характеристикой для проведения в ней очередного испытания.
-Это испытание либо вычисляет значение целевой функции $\varphi(y)$ (если выбранная подзадача принадлежала уровню $j=M$), либо порождает новые подзадачи согласно (\ref{block_nested_i}) при $j\leq M-1$.
-В последнем случае новые порожденные задачи добавляются к текущему множеству задач, вычисляются их характеристики и процесс повторяется.
-Завершение процесса оптимизации происходит, когда в корневой задаче выполняется условие остановки алгоритма, решающего эту задачу.
+Let us assume the nested subproblems (\ref{block_nested_i}) to be solved with the use of a multidimensional 
+global search algorithm described in Section \ref{SectionPeano}. Then, each subproblem (\ref{block_nested_i}) can be 
+associated with a numerical value called the characteristic of this problem. The value $R(t)$ from (\ref{MaxR}) (i.e., the 
+maximum characteristic of the intervals formed within the subproblem) can be taken as such characteristic. According to the 
+rule of computing the characteristics (\ref{R}), the higher the value of the characteristic, the more promising the 
+subproblem for continuing the search of the global minimum of the initial problem (\ref{main_problem}). Therefore, the 
+subproblem with the highest characteristic is selected for executing the next trial at each iteration. This trial either 
+computes the objective function value $\varphi(y)$ (if the selected subproblem belongs to the level $j=M$) or generates 
+new subproblems according to (\ref{block_nested_i}) when $j\leq M-1$. In the latter case, new generated problems are 
+added to current problem set, their characteristics are computed, and the process is repeated. The  optimization process is 
+finished when the  stop condition is fulfilled for the algorithm solving the root problem.
 
@@
 
-One of the well-known approaches to investigating and comparing multiextremal optimization algorithms is based on the application of these methods for solving a set of test problems generated randomly.
+One of the well-known approaches to investigating and comparing multiextremal 
+optimization algorithms is based on the application of these methods for 
+solving a set of test problems generated randomly.
+The comparison of the algorithms has been carried out using the Grishagin test problems $F_{gr}$ (see, for example, \cite{Grishagin1994}, test function 4) and the GKLS generator \cite{Gaviano2003}.
 
-The numerical comparison of the algorithms has been carried out using the Grishagin test problems $F_{gr}$ (описание данных функций см., например, в \cite{Grishagin1994}, test function 4) and the GKLS generator \cite{Gaviano2003}.
+In \cite{Barkalov2015,Grishagin2018} the global search algorithm (GSA) with the use of the evolvents as well as in combination 
+with the adaptive dimensionality reduction scheme was shown to overcome many well known optimization 
+algorithms including DIRECT \cite{Jones1993} and DIRECT\textit{l} \cite{Gablonsky2001}.
+Therefore, in the present study we will limit ourselves to the comparison of the  variants of GSA with different dimensionality reduction schemes.
 
-В работах \cite{Barkalov2015,Grishagin2018} было показано, что global search algorithm как с использованием разверток, так и в сочетании с адаптивной схемой редукции размерности превосходит многие известные алгоритмы, включая методы DIRECT и DIRECT\textit{l}.
-Поэтому в данном исследовании мы ограничимся сравнением вариантов GSA с различными схемами редукции размерности.
+In order to compare the efficiencies of the algorithms, the two criteria were used: the average number of trials and the 
+operating characteristic.
+The operating  characteristic of an algorithm is the function $P(k)$ defined as the fraction of the problems from the 
+considered series, for solving of which not more than $k$ trials have been required.
+The problem was considered to be solved, if the algorithm generated a trial point $y^k$ in the vicinity of 
+the global minimizer $y^\ast$, i.e. $\left|y^k-y^\ast\right| < \delta \left\|b-a\right\|$, where $\delta = 0.01$, $a$ and $b$ are the 
+boundaries of the search domain $D$.
 
-Для сравнения эффективности работы алгоритмов используются два критерия: the average number of trials and operational characteristic.
-Операционной характеристикой алгоритма называется функция $P(k)$, определяемая как доля задач, для решения которых потребовалось не более чем $k$ испытаний.
-The global minimizer $y^\ast$ was considered to be found, if the algorithm generated a trial point $y^k$ in the vicinity of the global minimum, i.e.
-$\left|y^k-y^\ast\right| < \delta \left\|b-a\right\|$, where $\delta = 0.01$, $a$ and $b$ are the boundaries of the search domain $D$.
-
-Первая серия экспериментов была проведена на двумерных задачах классов $F_{gr}$, GKLS \textit{Simple}, GKLS \textit{Hard} (100 функций каждого класса).
-В таблице 1  представлено среднее число испытаний, выполненных GSA с использованием evolvents ($K_e$), nested optimization scheme ($K_n$ ), adaptive nested optimization scheme ($K_a$).
+The first series of experiments has been carried out on the two-dimensional problems from the classes $F_{gr}$, GKLS 
+\textit{Simple}, and GKLS \textit{Hard} (100 functions from each class). Table~\ref{tab1} presents the averaged 
+numbers of trials executed by GSA with the use of evolvents ($K_e$), nested optimization scheme ($K_n$), and adaptive 
+nested optimization scheme ($K_a$). Fig.~\ref{fig1} and Fig.~\ref{fig2}(a,b) present the operating characteristics of the 
+algorithms obtained on the problem classes $F_{gr}$, GKLS \textit{Simple}, and GKLS \textit{Hard} respectively. The solid 
+line corresponds to the algorithm using the evolvents, short dashed line -- to the adaptive nested optimization scheme, and 
+the long dashed line -- to the nested optimization scheme. The results of experiments demonstrate that GSA with the use of 
+the adaptive nested optimization scheme shows almost the same speed as compared to GSA with the evolvents, and 
+both algorithms considerably exceed the algorithm using the nested optimization scheme. Therefore further experiments 
+were limited to the comparison of different variants of the adaptive  dimensionality reduction scheme.
 
@@
 \centering
-\caption{Average number of iterations for 2D problems.}\label{tab1}
+\caption{Average number of trials for 2D problems.}\label{tab1}
 \begin{tabular}{lccc}
 \hline\noalign{\smallskip}
- &  $F_{gr}$  &  GKLS \textit{Simple} &  GKLS \textit{Hard} \\
+ &  $\;\;\;F_{gr}\;\;\;$  &  GKLS \textit{Simple} &  GKLS \textit{Hard} \\
 \noalign{\smallskip}\hline\noalign{\smallskip}
@@
 
-Результаты экспериментов показывают, что GSA с использованием adaptive nested optimization scheme не сильно уступает по быстродействию GSA с развертками, и оба они значительно превосходят GSA в сочетании с nested optimization scheme.
+\begin{figure}
+\centering
+\includegraphics[width=0.50\textwidth]{2D.pdf}
+\caption{Operating characteristics using $F_{gr}$ class} 
+\label{fig1}
+\end{figure}
+
+\begin{figure}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{2DSimple.pdf} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{2DHard.pdf} \\ (b)}
+\end{minipage}
+\caption{Operating characteristics using 2d GKLS \textit{Simple} (a) and \textit{Hard} (b) classes}
+\label{fig2}
+\end{figure}
+
+The second series of experiments has been carried out on the four-dimensional problems from the classes GKLS 
+\textit{Simple} and GKLS \textit{Hard} (100 functions of each class). Table~\ref{tab2}  presents the averaged numbers 
+of trials executed by GSA with the use of the adaptive nested optimization scheme ($K_a$) and block adaptive nested 
+optimization scheme ($K_{ba}$) with two levels of subproblems of equal dimensionality $N_1=N_2=2$. 
+Note that when solving the problem of the dimensionality $N=4$ using the initial variant of the adaptive scheme, four 
+levels of one-dimensional subproblems are formed that complicates the processing of these ones.
+
+Fig.~\ref{fig3}(a,b) presents the operating characteristics of the algorithms obtained on the classes GKLS \textit{Simple} 
+and GKLS \textit{Hard} respectively. The dashed line corresponds to GSA using the adaptive nested optimization 
+scheme, the solid line -- the block adaptive nested optimization scheme. 
+The results of experiments demonstrate the use of the block adaptive nested optimization scheme provides a considerable 
+gain in the number of trials (up to 35\%) as compared to the initial adaptive nested optimization scheme.
+
+\begin{table}
+\centering
+\caption{Average number of trials for 4D problems.}\label{tab2}
+\begin{tabular}{lcc}
+\hline\noalign{\smallskip}
+ &    GKLS \textit{Simple} &  GKLS \textit{Hard} \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+ $K_a$  &  21747 & 35633 \\
+ $K_{ba}$ &  13894 & 31620 \\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
+\begin{figure}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{4DSimple.pdf} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{4DHard.pdf} \\ (b)}
+\end{minipage}
+\caption{Operating characteristics using 4d GKLS \textit{Simple} (a) and \textit{Hard} (b) classes}
+\label{fig3}
+\end{figure}
 
@@
 
-В данной работе предложена обобщенная адаптивная схема редукции размерности для задач глобальной оптимизации, комбинирующая использование кривых Пеано и схему вложенной оптимизации.
-Для решения редуцированных подзадач меньшей размерности используется алгоритм глобального поиска.
-Приведена вычислительная схема алгоритма, рассмотрены алгоритмические моменты работы адаптивной схемы редукции размерности.
-Проведены вычислительные эксперименты на серии тестовых задач с целью сравнения эффективности различных схем редукции размерности. 
-Результаты экспериментов показывают, что . 
-Перспективы ...
+In the present work, the generalized adaptive dimensionality reduction scheme for the global optimization problems 
+combining the use of Peano space-filling curves and the nested (recursive) optimization scheme has been proposed.
+For solving the reduced subproblems of less dimensionality, the global search algorithm was applied. 
+The computational scheme of the algorithm was given, main issues related to the use of the adaptive dimensionality 
+reduction scheme were considered.
+The numerical experiments have been carried out using the series of test problems in order to compare the efficiencies of 
+different dimensionality reduction schemes.  
+The result of experiments demonstrated that the use of the block adaptive nested optimization scheme can essentially reduce the 
+number of trials required to solve the problem with given accuracy. 
+Further works on the development of the global optimization methods based on this method of the 
+dimensionality reduction may be related to the use of the local estimates of the Lipschitz constant (considered, for example, in \cite{Grishagin2015,Sergeyev2016}) in different subproblems.
 

### 2019-01-27 2019/WCGO/WCGO.tex
New title

@@
 %
-\title{Generalized dimensionality reduction scheme for global optimization 
-algorithms
-\thanks{This study was supported by the Russian Science Foundation, project 
-No.\,16-11-10150.}}
+\title{Adaptive global optimization based on nested dimensionality 
+reduction\thanks{This study was supported by the Russian Science Foundation, 
+project No.\,16-11-10150.}}
 %
-\titlerunning{Generalized dimensionality reduction scheme}
+\titlerunning{Adaptive global optimization}
 % If the paper title is too long for the running head, you can set 

### 2019-02-11 2019/MOTOR_Sovrasov/motor_paper.tex
MOTOR: update the paper

@@
 \usepackage{amssymb}
+\usepackage{amsmath}
 \usepackage{subfig}
 \usepackage{cite}
+
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator{\sign}{sign}
+
+%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
+\usepackage[utf8x]{inputenc}
+\usepackage[english,russian]{babel}
+\usepackage{cmap}
 
@@
 %
-\title{Comparison of several derivative-free global optimization algorithms\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}}
+\title{Comparison of several stochastic and deterministic derivative-free global optimization algorithms\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}}
 %
@@
 \begin{abstract}
-This work considers several stochastic and deterministic deri-vative-free global optimization
-algorithms. In the first part of the paper popular sequential open-source solvers are compared against
-Globalizer solver, which is developed at the Lobachevsky State University. The Globalizer is
-designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows
+In this paper popular sequential open-source solvers are compared against
+Globalizer solver, which is developed at the Lobachevsky State University.
+The Globalizer is designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows
 competitive performance with other similar solvers. The comparison is done on several sets of
-challenging multi-extremal benchmark functions. The second part of this work is devoted to
-comparison between the Globalizer and MIDACO solvers on systems with shared and distributed
-memory. MIDACO is a state-of-the-art global solver included to the TOMLAB optimization
-environment for MATLAB. Results of the benchmark show advantages of the Globalizer on small-dimensional, but sufficiently multi-extremal benchmark functions.
+challenging multi-extremal benchmark functions. Also this work considers a method of
+heuristic hyperparameters control for Globalizer allowing to reduce amount of
+initial tuning before optimization. The proposed scheme allows rapidly increase convergence speed
+of Globalizer by switching between ``local'' and ``global'' search phases in runtime.
 % We would like to encourage you to list your keywords within
@@
 
-Nonlinear global optimization of non-convex functions is considered to be one of the most
-difficult problems of mathematical programming. The finding of the global minimum of
-a function of several variables often appears to be more difficult than local optimization in a
-thousand-dimensional space. In local optimization applying of the simplest gradient descent
-method may appear to be enough, whereas in order to \textit{guarantee} finding the global
-optimum, an optimization method has to accumulate information about behavior of the
-objective function within the whole search domain
-\cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}. Nowadays, various stochastic global
-optimization algorithms have attracted much attention, first of all, the evolution ones
-\cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}. They have rather simple structure
-and allow solving the problems of large dimensionality. However, the evolution algorithms ensure
-the global convergence in the probabilistic sense only.
+Задача отысканя глобального минимума нелинейных невыпуклых функций традиционно считается одной из самых трудных
+задач математического программирования. Зачастую она оказывается сложнее, чем локальная оптимизация в существенно многомерном пространстве.
+Для последней может оказаться достаточно применения простейшего метода градиентного спуска или pattern search algorithms \cite{torczon1997},
+в то время как чтобы \textit{гаранитрованно} отыскать глобальный оптимум методам
+оптимизации приходится накапливать информацию о поведении целевой функции во всей области поиска \cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}. В последнее время стали популярны
+различные стохастические алгоритмы глобально оптимизации, прежде всего эволюционные \cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}.
+Они имеют довольно простую структуру, позволяют решать задачи большой размерности, однако обеспечивают глобальную сходимость только в вероятностном смысле.
 
-In the present paper, the open-source implementations of nine different global optimization methods
-were considered. All the algorithms were tested using a set of 900 essentially multiextremal functions,
-which has been generated using generators of test problems \cite{Gaviano2003, grishaginClass}.
-Besides the comparison of the sequential algorithms, a comparison of the MIDACO solver
-\cite{Schlueter2012} with the Globalizer software system \cite{globalizerSystem,Strongin2018} on a subset of 200 test problems has been
-performed in the Lobachevsky supercomputer environment.
+В данной работе рассмотрены open-source реализации восьми различных методов глобальной оптимизации, представленные
+в библиотеке NLOpt\cite{nlopt} и пакете SciPY\cite{scipy}.
+Все алгоритмы были протестированы на наборе из 900 существенно многоэкстремальных функций, который был сгенерирован с
+помощью специализированных генераторов задач \cite{Gaviano2003, grishaginClass}.
 
@@
 present work. Using these ones, 9 sets of 100 problems of various complexity with the
-dimensionality varying from 2 to 5 were generated. Besides the comparison of the sequential
-methods, a comparison of the efficiencies of two parallel algorithms is presented in the work as well.
+dimensionality varying from 2 to 5 were generated.
 
@@
 
-\subsection{Parallel Algorithm of Global Search}
+\subsection{Algorithm of Global Search}
+\label{sub:ags}
 \subsubsection{Dimension Reduction with Space-Filling Curves}
@@
 
-\subsubsection{Algorithm of Global Search on Shared Memory}
-\label{sub:ags}
-Parallel optimization methods applied in Globalizer to solve the reduced problem
-(\ref{eq:oneDimTask}) are based on the MAGS method, which can be presented as follows ---
+Optimization method applied in Globalizer to solve the reduced problem
+(\ref{eq:oneDimTask}) is based on the AGS method, which can be presented as follows ---
 see \cite{strongin1978}, \cite{strSergGO}.
+\par
+The algorithm considered for solving the stated problem implies generating
+a sequence of points \(x_k\), in which the values of the minimized function \(z_k = f(x_k)\)
+are computed. Let us call the process of computating the function value
+(including calculating an image \(y^k=y(x^k)\)) a trial, and the pair \((x^k,z^k)\) ---
+the result of the trial. A set of the pairs \(\{(x^k,z^k)\}, 1\leqslant k\leqslant n\)
+makes up the search information accumulated by the method after executing \(n\) steps.
 \par
@@
 
-\textit{Rule 1}. Renumber the points of the preceding trials by the lower indices in order of
-increasing value of coordinates
-$0=x_0<x_1<...<x_{k+1}=1$.
+\textit{Step 1.} Renumber the points in the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\),
+which includes the boundary points of the interval \([0,1]\) as well as the points of
+preceding trials, by the lower indices in order of increasing coordinate values  i.e.
+\begin{displaymath}
+0=x_0<x_1<\dotsc<x_{k+1}=1
+\end{displaymath}
+\par
+\textit{Step 2.} Assuming \(z_i=f(x_i),1\leqslant i\leqslant k\), compute the values
+\begin{equation}
+\label{eq:step2}
+\mu=\max_{1\leqslant i\leqslant k}\dfrac{|z_i-z_{i-1}|}{\Delta_i},
+\begin{matrix}
+    M =
+    \left\{
+    \begin{matrix}
+    r\mu,\mu>0 \\
+    1,\mu=0
+    \end{matrix} \right.
+    \end{matrix}
+\end{equation}
+where \(r>1\) is a predefined parameter for the method, and \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
+\par
+\textit{Step 3.} For each interval \((x_{i-1},x_i),1\leqslant i\leqslant k+1\), compute the
+characteristics according to the formulae
+\begin{equation}
+\label{step3_1}
+R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
+\end{equation}
+\begin{equation}
+\label{eq:step3_2}
+R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1.
+\end{equation}
+\par
+\textit{Step 4.} Determine the interval with the maximum characteristic \((x_{t-1}, x_t),\:t=\argmax_{1\leqslant i \leqslant k+1}R(i)\)
 
-\textit{Rule 2}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq
-k+1\).
-
-\textit{Rule 3}. Determine the \(p\) intervals with the maximum characteristics $R(t_j)=\max_{1\leq i
-\leq k+1}R(i),\: j=\overline{1,p}$.
-
-\textit{Rule 4}. Execute new trials at points \(x^{k+j},\: j=\overline{1,p}\) located within intervals
-with
-maximum characteristics from the previous step
-  $x^{k+j}=d(x_{t_j}),\: j=\overline{1,p}$.
+\textit{Step 5.} Execute a new trial at point \(x_{k+1}\) computed according to the formula
+\begin{displaymath}
+x_{k+1}=\dfrac{x_{t}+x_{t-1}}{2},t=1,t=k+1,
+\end{displaymath}
+\begin{displaymath}
+\label{step5}
+x_{k+1}=\dfrac{x_{t}+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{\mu}\right]^N,1<t<k+1.
+\end{displaymath}
 
 The stopping condition, which terminated the trials, is defined by the inequality
-$\rho_{t_j}<\varepsilon,\: j=\overline{1,p}$
-for the intervals with maximum characteristics from Step 3 and \(\varepsilon >0\) is the
+\(\Delta_{t}\leqslant \varepsilon\)
+for the interval with the maximum characteristics from Step 4 and \(\varepsilon >0\) is the
 predefined accuracy of the optimization problem solution. If the stopping condition is not satisfied,
-the index \(k\) is incremented by \(p\), and the new global optimization iteration is executed.
+the index \(k\) is incremented by \(1\), and the new global optimization iteration is executed.
 
-This method is employed in Globalizer to organize parallel computations on shared memory: each of
-\(p\)
-trials can be carried out on one of \(p\) local computation units.
+The convergence conditions of the described algorithm are given, for example, in \cite{strSergGO}.
 
-The convergence conditions and exact formulas for decision rules $R(i)$ and $d(x)$ of the
-described algorithm are given, for example, in \cite{strSergGO}.
+\subsubsection{Hyperparameters tuning}
+Параметр $r$ из (\ref{eq:step2}) непосредственно влияет на глобальную сходимость AGS (see \cite{strSergGO}, Chapter 8):
+при достаточно большом значении $r$ метод гарантированно сходится ко всем глобальным минимумам целевой функции.
+В то же время, согласно (\ref{eq:step3_2}), при бесконечно большом значении $r$ AGS превращается в полный перебор
+по равномерной сетке.
 
-\subsection{Sequential Methods}
+В идеальном случае для обеспечения наибольшей скорости сходимости оценка константы Липшица из (\ref{eq:step2})
+не должна быть слишком завышенной, но на практике реальное значение $L$ из (\ref{eq:lip}) неизвестно и приходится
+либо брать заведомо большое значение $r$, либо проводить несколько запусков AGS с разными параметрами. Чтобы в некотрой
+степени решить обозначенную проблему выбора $r$ будем использовать следующую схему:
 \begin{itemize}
-  \item \textbf{Algorithm of Global Search}. Sequential version of the method described in Section
-\ref{sub:ags}.
+  \item совершить $q$ итераций AGS при $r=r_{max}$;
+  \item совершить $q$ итераций AGS при $r=r_{min}$;
+  \item повторять предыдущие шаги до схоимости или исчерпания лимита итараций.
+\end{itemize}
 
-  \item \textbf{Locally-biased Algorithm of Global Search (AGS\(l\))} \cite{indexMethod}. It's a
-modification of
-  the original AGS which make it more locally oriented by alternately using of two types of
-characteristics in the Rule 2 from the Section \ref{sub:ags}.
+В указанном алгоритме $r_{min} < r_{max}$, $q > 1$. Вместо одного параметра $r$ теперь
+необходимо выбирать 3, однако, согласно численным экспериментам, сделать это проще, чем найти оптимальное значение $r$.
+Интуитивно практическую эффективность предложенной схемы можно объяснить тем, что теперь работа метода происходит в
+двух режимах: глобальный поиск при $r=r_{max}$ и локальный при $r=r_{min}$. Если во время фазы глобального посика
+метод приблизился к глобальному минимуму, то во время следующей фазы оценка глобального минимума будет быстро уточнена.
+Если двух фаз не хватило, то процесс продолжается. Таким образом выбирается более оптимальный trade-off between
+exploration and exploitation. Далее будем обозначать метод, использующий описанную схему как AGS-AR.
 
+\subsection{Other Optimization Methods}
+\begin{itemize}
   \item \textbf{Multi Level Single Linkage} \cite{Kan1987StochasticGO}. MLSL is an improved
@@
 
-All the mentioned algorithms are available in source codes as parts of wide-spread optimization packages (except of AGS and AGS\(l\), which are implemented in the Globalizer system).
+All the mentioned algorithms are available in source codes as parts of wide-spread optimization packages.
 DIRECT, DIRECT$l$, CRS, MLSL and StoGO are part of the NLOpt library \cite{nlopt}.
@@
 \begin{itemize}
-  \item in the AGS\(l\) method, the parameter of alternation the
-global and local iterations was set to be equal to 5:1;
+  \item in the AGS-AR method, the parameter of alternation the
+  global and local stages $q$ was set to be equal to $100\cdot N^2$, also $r_{min}=3,\:r_{max}=2\cdot r_{min}$;
   \item in the DIRECT and DIRECT\(l\) methods, the parameter \(\epsilon=10^{-4}\);
@@
 The rest parameters were varied subject to the problem class (see Table \ref{tab:params}).
+For the AGS the value ot the $r$ parameter, such that the method solves all problems and performs the minimum amount of trials,
+was estimated by brute force on the uniform grid with step $0.1$.
 
@@
 The results of running the optimization methods on the considered problem classes are presented in
-Tables \ref{tab:trials}, \ref{tab:solved}. The DIRECT and AGS\(l\) methods have demonstrated the
-best convergence speed on all classes, at that AGS\(l\) inferior to DIRECT on the problems from the
+Tables \ref{tab:trials}, \ref{tab:solved}. The DIRECT, AGS and AGS-AR methods have demonstrated the
+best convergence speed on all classes, at that AGS-AR inferior to DIRECT on the 2d problems from the
 Simple classes and has an advantage on the problems of the Hard classes. As one can see from Table
-\ref{tab:solved}, the deterministic methods (AGS, AGS\(l\), DIRECT, and DIRECT\(l\)) were the
+\ref{tab:solved}, the deterministic methods (AGS, AGS-AR, DIRECT, and DIRECT\(l\)) were the
 most reliable. Among the stochastic methods, MLSL and SDA have demonstrated the highest
@@
 \begin{center}
-\caption{Averaged number of trials executed by sequential methods for solving the test
+\caption{Averaged number of trials executed by optimization methods for solving the test
 optimization problems}
@@
     \hline
-    & AGS & AGS\(l\) & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
+    & AGS & AGS-AR & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
   \hline
@@
 \begin{center}
-\caption{Number of test optimization problems solved by sequential methods}
+\caption{Number of test optimization problems solved by methods}
   \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
     \hline
-    & AGS & AGS\(l\) & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
+    & AGS & AGS-AR & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
   \hline
@@
 \begin{itemize}
-  \item AGS\(l\) method implemented in the Globalizer system has demonstrated the convergence
+  \item the proposed modification of the stock AGS substantially outperforms it on the most complicated problems
+  from the generated test set and allows to pay less attention to initial hyperparameter tuning.
+  \item AGS-AR method  has demonstrated the convergence
 speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-source
@@
 problems;
-  \item the parallel version of the Globalizer system demonstrates good speedup
-when running on several nodes, on each of which a single objective function value per iterations is
-computed. When solving the problems with fast computable objective functions and using multiple
-threads on the nodes, the speedup for the Globalizer system degrade with increasing
-number of nodes;
-  \item the MIDACO system is the most suitable for simple global optimization problems with the
-fast-computable objective functions. In this case, MIDACO is reliable enough and provides a linear
-speedup with increasing number of nodes and threads executed on these ones in parallel.
 \end{itemize}

### 2019-02-11 2019/MOTOR_Sovrasov/motor_paper.tex
MOTOR: update conclusions

@@
 \section{Introduction}
+\begin{Russian}
 
@@
 помощью специализированных генераторов задач \cite{Gaviano2003, grishaginClass}.
+\end{Russian}
 
@@
 \subsubsection{Hyperparameters tuning}
+
+\begin{Russian}
 Параметр $r$ из (\ref{eq:step2}) непосредственно влияет на глобальную сходимость AGS (see \cite{strSergGO}, Chapter 8):
@@
 exploration and exploitation. Далее будем обозначать метод, использующий описанную схему как AGS-AR.
+\end{Russian}
 
@@
   \hline
-  GKLS 2d Hard   &  100 & 100 & 74  & 100 & 100 & 100 & 93  & 85  & 77 \\
+  GKLS 2d Hard   &  100 & 97 & 74  & 100 & 100 & 100 & 93  & 85  & 77 \\
   \hline
-  GKLS 3d Simple &  100 & 97  & 75  & 100 & 100 & 100 & 89  & 86  & 44 \\
+  GKLS 3d Simple &  100 & 100  & 75  & 100 & 100 & 100 & 89  & 86  & 44 \\
   \hline
-  GKLS 3d Hard   &  100  & 99   & 72   & 100  & 99   & 100  & 88   & 77   & 43 \\
+  GKLS 3d Hard   &  100  & 100   & 72   & 100  & 99   & 100  & 88   & 77   & 43 \\
   \hline
@@
   \hline
-  GKLS 5d Hard   &  97  & 99  & 42  & 100 & 90  & 79  & 84  & 48  & 8 \\
+  GKLS 5d Hard   &  97  & 100  & 42  & 100 & 90  & 79  & 84  & 48  & 8 \\
   \hline
@@
 \begin{itemize}
-  \item the proposed modification of the stock AGS substantially outperforms it on the most complicated problems
-  from the generated test set and allows to pay less attention to initial hyperparameter tuning.
+  \item the proposed modification of the stock AGS, AGS-AR allows to pay less attention to initial hyperparameter tuning and
+  performs on-par with properly tuned AGS.
   \item AGS-AR method has demonstrated the convergence

### 2019-02-13 2019/MOTOR_Sovrasov/motor_paper.tex
MOTOR: update paper

@@
 \DeclareMathOperator{\sign}{sign}
-
-%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
-\usepackage[utf8x]{inputenc}
-\usepackage[english,russian]{babel}
-\usepackage{cmap}
 
@@
 \section{Introduction}
-\begin{Russian}
 
-Задача отысканя глобального минимума нелинейных невыпуклых функций традиционно считается одной из самых трудных
-задач математического программирования. Зачастую она оказывается сложнее, чем локальная оптимизация в существенно многомерном пространстве.
-Для последней может оказаться достаточно применения простейшего метода градиентного спуска или pattern search algorithms \cite{torczon1997},
-в то время как чтобы \textit{гаранитрованно} отыскать глобальный оптимум методам
-оптимизации приходится накапливать информацию о поведении целевой функции во всей области поиска \cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}. В последнее время стали популярны
-различные стохастические алгоритмы глобально оптимизации, прежде всего эволюционные \cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}.
-Они имеют довольно простую структуру, позволяют решать задачи большой размерности, однако обеспечивают глобальную сходимость только в вероятностном смысле.
+The problem of finding the global minima of the nonlinear nonconvex functions is considered to be
+one of the most difficult mathematical programming problems traditionally. Often, it appears to be
+more complex than the local optimization in an essentially multidimensional space.
+For the latter, the application of the simplest gradient descent method or of the pattern search
+algorithms may appear to be sufficient \cite{torczon1997}
+whereas in order to \textit{guarantee} the finding of the global optimum, the optimization methods
+have to
+accumulate the information on the behavior of the objective function in the whole search domain
+\cite{Jones2009,Paulavicius2011,Evtushenko2013,strSergGO}.
+Recently, various stochastic global optimization algorithms, first of all, the evolution ones
+\cite{Storn1997, SCHLUTER2009, KennedyEberhart1995} became popular. These ones have rather
+simple structure and allow solving the problems of large dimensionality. However, these methods
+provide the global convergence in the probabilistic meaning only.
 
-В данной работе рассмотрены open-source реализации восьми различных методов глобальной оптимизации, представленные
-в библиотеке NLOpt\cite{nlopt} и пакете SciPY\cite{scipy}.
-Все алгоритмы были протестированы на наборе из 900 существенно многоэкстремальных функций, который был сгенерирован с
-помощью специализированных генераторов задач \cite{Gaviano2003, grishaginClass}.
-\end{Russian}
+In the present work, the open-source implementations of the eight different global optimization
+methods included into the NLOpt library \cite{nlopt} and the SciPY package \cite{scipy} are
+considered.
+All algorithms were tested on a set of 900 essentially multiextremal functions, which has been
+generated with the use of special problem generators \cite{Gaviano2003, grishaginClass}.
 
@@
 
-\begin{Russian}
-Параметр $r$ из (\ref{eq:step2}) непосредственно влияет на глобальную сходимость AGS (see \cite{strSergGO}, Chapter 8):
-при достаточно большом значении $r$ метод гарантированно сходится ко всем глобальным минимумам целевой функции.
-В то же время, согласно (\ref{eq:step3_2}), при бесконечно большом значении $r$ AGS превращается в полный перебор
-по равномерной сетке.
+The parameter $r$ from (\ref{eq:step2}) affects the global convergence of AGS directly (see
+\cite{strSergGO}, Chapter 8):
+at high enough value of $r$, the method converges to all global minima of the objective function with
+guarantee.
+At the same time, according to (\ref{eq:step3_2}), at the infinitely high value of $r$, AGS turns into
+the brute force search method on a uniform grid.
 
-В идеальном случае для обеспечения наибольшей скорости сходимости оценка константы Липшица из (\ref{eq:step2})
-не должна быть слишком завышенной, но на практике реальное значение $L$ из (\ref{eq:lip}) неизвестно и приходится
-либо брать заведомо большое значение $r$, либо проводить несколько запусков AGS с разными параметрами. Чтобы в некотрой
-степени решить обозначенную проблему выбора $r$ будем использовать следующую схему:
+In the ideal case, in order to provide the highest convergence speed, the estimate of the Lipschitz
+constant from (\ref{eq:step2})
+should not be too overestimated, but in practice the actual value of $L$ from (\ref{eq:lip}) in
+unknown, and one has either to take an obviously overestimated value of $r$ or to execute several
+runs of AGS with different parameters. In order to resolve the problem of choosing $r$ to some extent,
+let us use the following scheme:
 \begin{itemize}
-  \item совершить $q$ итераций AGS при $r=r_{max}$;
-  \item совершить $q$ итераций AGS при $r=r_{min}$;
-  \item повторять предыдущие шаги до схоимости или исчерпания лимита итараций.
+  \item execute $q$ iterations of AGS with $r=r_{max}$;
+  \item execute $q$ iterations of AGS with $r=r_{min}$;
+  \item repeat the above steps either until convergence or until the allowed number of iterations are
+exhausted.
 \end{itemize}
 
-В указанном алгоритме $r_{min} < r_{max}$, $q > 1$. Вместо одного параметра $r$ теперь
-необходимо выбирать 3, однако, согласно численным экспериментам, сделать это проще, чем найти оптимальное значение $r$.
-Интуитивно практическую эффективность предложенной схемы можно объяснить тем, что теперь работа метода происходит в
-двух режимах: глобальный поиск при $r=r_{max}$ и локальный при $r=r_{min}$. Если во время фазы глобального посика
-метод приблизился к глобальному минимуму, то во время следующей фазы оценка глобального минимума будет быстро уточнена.
-Если двух фаз не хватило, то процесс продолжается. Таким образом выбирается более оптимальный trade-off between
-exploration and exploitation. Далее будем обозначать метод, использующий описанную схему как AGS-AR.
-\end{Russian}
+In the above algorithm, $r_{min} < r_{max}$, $q > 1$. Instead one parameter $r$, now
+3 ones should be selected. However, according to the results of the numerical experiments, it is easier
+than to find the optimal value of $r$.
+Intuitively, the practical efficiency of the proposed scheme can be explained by the fact that now the
+operation of the method takes place in two modes: the global search with $r=r_{max}$ and the local
+one with $r=r_{min}$. If during the global search phase, the method approached the global minimum
+whereas during the next phase, the estimate of the global minimum  would be refined rapidly.
+If two phases are not enough, the process is continued. This way, a better trade-off
+between the exploration and the exploitation is achieved.
+Further, we will denote the method utilizing the scheme described above as AGS-AR.
 
@@
 \(y^k\) in a \(\delta\)-vicinity of the global minimum \(y^*\), i.e. $\left\|y^k-y^*\right\|\leq \delta
-= 0.01\left\|b-a\right\|$, where \(a\) and \(b\) are the left and the right boundaries of the hypercube
-from (\ref{eq:task}). If this relation is not fulfilled before the expiration of the limit of the number of
-trials, the problem was considered to be unsolved. The limit of the number of trials was set
+= \alpha\left\|b-a\right\|$, where \(a\) and \(b\) are the left and the right boundaries of the hypercube
+from (\ref{eq:task}), $\alpha$ is relative precision. If this relation is not fulfilled before the expiration of the limit of the number of
+trials, the problem was considered to be unsolved. The limit of the number of trials and $\alpha$ were set
 for each problem class according to the dimensionality and complexity (see Table \ref{tab:limits}).
@@
 \begin{center}
-\caption{Trials limits for the test problem classes}
-  \begin{tabular}{|l|{c}|}
+\caption{Trials limits and relative precision for the test problem classes}
+  \begin{tabular}{|l|{c}|{c}|}
     \hline
-  Problems class & Trials limit\\
+  Problems class & Trials limit & $\alpha$\\
   \hline
-  \(F_{GR}\) & 5000 \\
+  \(F_{GR}\) & 5000 & 0.01 \\
   \hline
-  GKLS 2d Simple & 8000 \\
+  GKLS 2d Simple & 8000 & 0.01 \\
   \hline
-  GKLS 2d Hard & 9000 \\
+  GKLS 2d Hard & 9000 & 0.01 \\
   \hline
-  GKLS 3d Simple & 15000 \\
+  GKLS 3d Simple & 15000 & 0.01 \\
   \hline
-  GKLS 3d Hard & 25000 \\
+  GKLS 3d Hard & 25000 & 0.01 \\
   \hline
-  GKLS 4d Simple & 150000 \\
+  GKLS 4d Simple & 150000 & $\sqrt[4]{10^{-6}}$ \\
   \hline
-  GKLS 4d Hard & 250000 \\
+  GKLS 4d Hard & 250000 & $\sqrt[4]{10^{-6}}$ \\
   \hline
-  GKLS 5d Simple & 350000 \\
+  GKLS 5d Simple & 350000 & $\sqrt[5]{10^{-7}}$ \\
   \hline
-  GKLS 5d Hard & 600000 \\
+  GKLS 5d Hard & 600000 & $\sqrt[5]{10^{-7}}$ \\
   \hline
@@
   \item in the AGS-AR method, the parameter of alternation the
-  global and local stages $q$ was set to be equal to $100\cdot N^2$, also $r_{min}=3,\:r_{max}=2\cdot r_{min}$;
+  global and local stages $q$ was set to be equal to $50\cdot\log_2(N-1)\cdot N^2$, also $r_{min}=3,\:r_{max}=2\cdot r_{min}$;
   \item in the DIRECT and DIRECT\(l\) methods, the parameter \(\epsilon=10^{-4}\);
@@
   \hline
-  \(F_{GR}\)     & 193.1 &  \textbf{158.3} & 400.3 & 182.3 & 214.9 & 947.2 & 691.2 & 1257.3 &
-1336.8 \\
+  \(F_{GR}\)     & 193.1 & 248.3 & 400.3 & \textbf{182.2} & 214.9 & 947.2 & 691.2 & 1257.3 & 1336.8 \\
   \hline
-  GKLS 2d Simple &  254.9 & 217.6 & 510.6 & \textbf{189.0} & 255.2 & 556.8 & 356.3 & 952.2
-& 1251.5 \\
+  GKLS 2d Simple & 254.9 & 221.6 & 510.6 & \textbf{189.0} & 255.2 & 556.8 & 356.3 & 952.2 & 1251.5 \\
   \hline
-  GKLS 2d Hard   &  728.7 & \textbf{488.0} & 844.7 & 985.4 & 1126.7 & 1042.5 & 1637.9 &
-1041.1 & 2532.2 \\
+  GKLS 2d Hard   & \textbf{728.7} & 785.0 & 844.7 & 985.4 & 1126.7 & 1042.5 & 1637.9 & 1041.1 & 2532.2 \\
   \hline
-  GKLS 3d Simple &  1372.1 & 1195.3 & 4145.8 & \textbf{973.6} & 1477.8 & 4609.2 & 2706.5 &
-5956.94 & 3856.1 \\
+  GKLS 3d Simple &  1372.1 & 1169.5 & 4145.8 & \textbf{973.6} & 1477.8 & 4609.2 & 2706.5 & 5956.9 & 3856.1 \\
   \hline
-  GKLS 3d Hard   &  3636.1 & \textbf{1930.5} & 6787.0 & 2298.7 & 3553.3 & 5640.1 & 4708.4 &
-6914.3 & 7843.2 \\
+  GKLS 3d Hard   &  3636.1 & \textbf{1952.1} & 6787.0 & 2298.7 & 3553.3 & 5640.1 & 4708.4 & 6914.3 & 7843.2 \\
   \hline
-  GKLS 4d Simple &  26654.1 & 11095.7 & 37436.8 & \textbf{7824.3} & 15994.1 & 41514.3 &
-21417.9 & 19157.7 & 59895.4 \\
+  GKLS 4d Simple &  5729.8 & \textbf{4919.1} & 19883.6 & 7328.8 & 15010.0 & 41484.8 & 22066.0 & 6271.2 & 29359.2 \\
   \hline
-  GKLS 4d Hard   &  54536.8 &  \textbf{23167.8} &  73779.3 &  23204.4 &  54489.9 &  80247.2
-&  68815.5 &  27466.1 &  109328.1  \\
+  GKLS 4d Hard   &  13113.4 & \textbf{12860.1} & 27137.4 & 22884.4 & 55596.1 & 80220.1 & 68048.0 & 12487.6 & 58925.5  \\
   \hline
-  GKLS 5d Simple &  29810.0 & 11529.0 & 143575.0 & \textbf{7166.5} & 13970.5 & 52647.6 &
-34255.3 & 73074.5 & 91580.4 \\
+  GKLS 5d Simple &  \textbf{5821.5} & 6241.3 & 62921.7 & 5966.1 & 10795.5 & 52609.2 & 34208.8 & 20859.4 & 69206.8 \\
   \hline
-  GKLS 5d Hard   &  113129.1 & 67652.7 & 165192.8 & \textbf{66327.4} & 164390.6 & 138766.2
-& 116973.1 & 105496.9 & 155123.8 \\
+  GKLS 5d Hard   &  \textbf{17008.6} & 21555.1 & 87563.9 & 61657.3 & 148637.8 & 138011.8 & 115634.6 & 26850.0 & 141886.5 \\
   \hline
@@
   \hline
-  GKLS 4d Simple &  100 & 100 & 46  & 100 & 100 & 94  & 78  & 59  & 16 \\
+  GKLS 4d Simple &  100 & 100 & 74 & 100 & 100 & 94 & 82 & 68 & 72 \\
   \hline
-  GKLS 4d Hard   &  100 & 100 & 47  & 99  & 97  & 94  & 72  & 32  & 10  \\
+  GKLS 4d Hard   &  100 & 100 & 60 & 99 & 99 & 94 & 73 & 55 & 69  \\
   \hline
-  GKLS 5d Simple &  100 & 100 & 68  & 100 & 100 & 98  & 100 & 77  & 9  \\
+  GKLS 5d Simple &  100 & 100 & 86 & 100 & 100 & 98 & 100 & 88 & 82  \\
   \hline
-  GKLS 5d Hard   &  97  & 100  & 42  & 100 & 90  & 79  & 84  & 48  & 8 \\
+  GKLS 5d Hard   &  100 & 100 & 77 & 100 & 93 & 79 & 86 & 77 & 78 \\
   \hline

### 2019-02-13 2019/MOTOR_Sovrasov/motor_paper.tex
MOTOR: add OC plots

@@
 
+\begin{figure}[ht]
+  \centering
+  \subfloat[4d Simple]{{\includegraphics[width=.5\textwidth]{images/gklss4d.pdf}}\label{fig:s4d}}
+  \subfloat[4d Hard]{{\includegraphics[width=.5\textwidth]{images/gklsh4d.pdf}}\label{fig:h4d}}
+  \caption{Operating characteristics of the algorithms when solving problems from the GKLS 4d classes}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \subfloat[5d Simple]{{\includegraphics[width=.5\textwidth]{images/gklss5d.pdf}}\label{fig:s4d}}
+  \subfloat[5d Hard]{{\includegraphics[width=.5\textwidth]{images/gklsh5d.pdf}}\label{fig:h4d}}
+  \caption{Operating characteristics of the algorithms when solving problems from the GKLS 5d classes}
+\end{figure}
+
 \section{Conclusions}

### 2019-02-14 2019/MOTOR_Sovrasov/motor_paper.tex
MOToR: pre-submit version

--- 2019-02-13 2019/MOTOR_Sovrasov/motor_paper.tex
@@
 %
-\title{Comparison of several stochastic and deterministic derivative-free global optimization algorithms\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}}
+\title{Comparison of Several Stochastic and Deterministic Derivative-free Global Optimization Algorithms\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}}
 %
@@
 \begin{abstract}
-In this paper popular sequential open-source solvers are compared against
+In this paper popular open-source solvers are compared against
 Globalizer solver, which is developed at the Lobachevsky State University.
@@
 challenging multi-extremal benchmark functions. Also this work considers a method of
-heuristic hyperparameters control for Globalizer allowing to reduce amount of
-initial tuning before optimization. The proposed scheme allows rapidly increase convergence speed
-of Globalizer by switching between ``local'' and ``global'' search phases in runtime.
+heuristic hyperparameters control for the Globalizer allowing to reduce amount of
+initial tuning before optimization. The proposed scheme allows substantially increase convergence speed
+of the Globalizer by switching between ``local'' and ``global'' search phases in runtime.
 % We would like to encourage you to list your keywords within
@@
 \keywords{deterministic global optimization $\cdot$ stochastic global optimization
-  $\cdot$ parallel numerical methods $\cdot$ derivative-free algorithms $\cdot$ black-box
-optimization}
+  $\cdot$ algorithms comparison $\cdot$ derivative-free algorithms $\cdot$ black-box
+optimization $\cdot$ multi-extremal problems}
 \end{abstract}
@@
 other has been considered in the literature. In these works, most of modern methods have
-been studied in details, but the emphasis was made mainly on the sequential algorithms while the
-reliability and the convergence speed were the main criteria of efficiency of the optimization
-methods. In the majority of works, the sets of well-known test problems (for example, the Rastrigin
+been studied in details. In the majority of works, the sets of well-known test problems (for example, the Rastrigin
 function, Ackley function, etc.) were taken as the sets of test functions. The sizes of such sets don't
@@
 
-Optimization method applied in Globalizer to solve the reduced problem
+Optimization method applied in Globalizer \cite{globalizerSystem} to solve the reduced problem
 (\ref{eq:oneDimTask}) is based on the AGS method, which can be presented as follows ---
@@
 \end{displaymath}
-\begin{displaymath}
+\begin{equation}
 \label{step5}
 x_{k+1}=\dfrac{x_{t}+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{\mu}\right]^N,1<t<k+1.
-\end{displaymath}
+\end{equation}
 
@@
 
-\subsubsection{Hyperparameters tuning}
+\subsubsection{Hyperparameters control in AGS}
 
@@
 guarantee.
-At the same time, according to (\ref{eq:step3_2}), at the infinitely high value of $r$, AGS turns into
+At the same time, according to (\ref{eq:step3_2}) and (\ref{step5}), at the infinitely high value of $r$, AGS turns into
 the brute force search method on a uniform grid.
@@
 \cite{Beiranvand2017}. In the present work, we will use two generators of test problems generating
-the problems of different nature \cite{grishaginClass, Gaviano2003}.
+the problems of different nature \cite{grishaginClass, Gaviano2003} \footnote{Software implementations of
+these generators are available in source codes at the page \url{https://github.com/sovrasov/global-optimization-test-problems}}.
 
@@
 
+The average number of trials doesn't represent the real behavior of an optimization method
+on a problems set in some cases. For an instance, if a method performs well on the most problems
+and spends too much trials to solve the least several problems, we wouldn't catch such
+case looking at the average number of trials only.
+As an advanced measure of performance we will use the operating characteristic \cite{grishaginClass}.
+It's defined by a set of points on the \((K, P)\) plane where \(K\) is the average number of search trials
+conducted before satisfying the termination condition when minimizing a function
+from a given class, and \(P\) is the proportion of problems solved successfully.
+If at a given \(K\), the operating characteristic of a method goes higher than one
+from another method, it means that at fixed search costs, the former method has a
+greater probability of finding the solution. If some value of \(P\) is fixed, and the
+characteristic of a method goes to the left from that of another method, the former
+method requires fewer resources to achieve the same reliability.
+
 \section{Results of Numerical Experiments}
-
+\label{sec:experiments}
 The results of various algorithms on different problem classes depend on the adjustments of
@@
 
-The results of running the optimization methods on the considered problem classes are presented in
-Tables \ref{tab:trials}, \ref{tab:solved}. The DIRECT, AGS and AGS-AR methods have demonstrated the
-best convergence speed on all classes, at that AGS-AR inferior to DIRECT on the 2d problems from the
-Simple classes and has an advantage on the problems of the Hard classes. As one can see from Table
-\ref{tab:solved}, the deterministic methods (AGS, AGS-AR, DIRECT, and DIRECT\(l\)) were the
-most reliable. Among the stochastic methods, MLSL and SDA have demonstrated the highest
-reliability.
-
 \begin{table}
@@
 
+The results of running the optimization methods on the considered problem classes are presented in
+Tables \ref{tab:trials}, \ref{tab:solved}. The DIRECT, AGS and AGS-AR methods have demonstrated the
+best convergence speed on all classes, at that AGS-AR inferior to DIRECT on the 2d problems from the
+Simple classes and has an advantage on the problems of the Hard classes. As one can see from Table
+\ref{tab:solved}, the deterministic methods (AGS, AGS-AR, DIRECT, and DIRECT\(l\)) were the
+most reliable. Among the stochastic methods, MLSL and SDA have demonstrated the highest
+reliability.
+
 \begin{table}
@@
 
+\begin{figure}[ht]
+  \centering
+  \subfloat[4d Simple]{{\includegraphics[width=.5\textwidth]{images/gklss4d.pdf}}\label{fig:s4d}}
+  \subfloat[4d Hard]{{\includegraphics[width=.5\textwidth]{images/gklsh4d.pdf}}\label{fig:h4d}}
+
+  \subfloat[5d Simple]{{\includegraphics[width=.5\textwidth]{images/gklss5d.pdf}}\label{fig:s5d}}
+  \subfloat[5d Hard]{{\includegraphics[width=.5\textwidth]{images/gklsh5d.pdf}}\label{fig:h5d}}
+  \caption{Operating characteristics of the algorithms when solving problems from the GKLS 4d and 5d classes. Best viewed in color.}
+\end{figure}
+
+Operating characteristic of the methods (Figures \ref{fig:s4d}, \ref{fig:h4d}, \ref{fig:s5d}, \ref{fig:h5d})
+demonstrates that AGS and AGS-AR faster than the other methods achieve 100\% success rate. Also on GKLS 5d Simple the DIRECT
+generally has the best performance, but there are several hard problems that affect it's average number of
+trials metric.
+
+\paragraph{Robustness of AGS and AGS-AR to the Hyperparameters Choice.}
+
+In order to investigate the influence of hyperparameters to the convergence speed of the AGS and AGS-AR,
+experiments with the following settings were conducted on the problems from GKLS 5d Simple class:
+\begin{itemize}
+  \item AGS with $r=4$ (like in the Table \ref{tab:params});
+  \item AGS with $r=6$;
+  \item AGS-AR with parameters from the beginning of the Section \ref{sec:experiments}
+  ($q=50\cdot\log_2(4)\cdot 25 = 2500$, $r_{min}=3,\:r_{max}=2\cdot r_{min}$);
+  \item AGS-AR with $r_{max}=8$ and other parameters from the previous experiment;
+  \item AGS-AR with $q=1000$ and other parameters from the beginning of the Section \ref{sec:experiments};
+\end{itemize}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.6\textwidth]{images/ar_stab.pdf}
+  \caption{Operating characteristics of AGS and AGS-AR with different hyperparameters
+  when solving problems from the GKLS 5d Simple classe. Best viewed in color.}
+  \label{fig:stability}
+\end{figure}
+
+The operating characteristics collected in the experiments described above are shown in the Figure \ref{fig:stability}.
+AGS with $r=6$ (the cyan-colored curve) shows the worst convergence speed, which indicates that AGS is very sensitive to choice of $r$.
+Since on the start AGS-AR has the same value of $r$ as AGS with $r=6$, operating characteristics of these methods
+are identical up to $K=2500$. After that point AGS-AR switches to $r=3$ and rapidly begins to increase the amount of solved problems
+until the next exploration phase on $K=5000$. Intervals where AGS-AR works with $r=r_{max}$ are visible on the operating characteristics as plateaus.
+Variations of $r$ and $q$ didn't drastically change the operating characteristic of AGS-AR. The latter observation shows robustness of the proposed
+AGS modification with the alternating parameter $r$.
+
 \section{Conclusions}
 
-In the present paper, several sequential and parallel global optimization algorithms were considered.
-A comparison of efficiencies of these ones has been done on a set of test problems. The results allow
-making the following conclusions:
+In the present paper, several global optimization algorithms were considered.
+A comparison of efficiencies of these ones has been done on a set of test problems.
+Also a scheme of hyperparameters control for the AGS algorithms was proposed and evaluated.
+The results presented in this work allow making the following conclusions:
 \begin{itemize}

### 2019-02-15 2019/MOTOR_Sovrasov/motor_paper.tex
MOTOR: fix captures of some tables
typography corrected

@@
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Comparison of Global Optimization Algorithms}
 % If the paper title is too long for the running head, you can set
@@
 Globalizer solver, which is developed at the Lobachevsky State University.
-The Globalizer is designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows
+The Globalizer is designed to solve problems with black-box objective function satisfying the Lipschitz condition and shows
 competitive performance with other similar solvers. The comparison is done on several sets of
@@
 
-In Ref. \cite{Beiranvand2017}, some general principles were formulated, which, in the author's
+In \cite{Beiranvand2017}, some general principles were formulated, which, in the author's
 opinion, should be obeyed when comparing the optimization methods. In particular, the authors say
@@
 generator can appear to be not enough for a comprehensive comparison of the methods. In order to
-overcome this problem in part, the authors of Ref. \cite{Beiranvand2017} advise to use several
+overcome this problem in part, the authors of the paper \cite{Beiranvand2017} advise to use several
 generators of various nature and to create the sets of problems of various complexity.
@@
     \hline
-    & AGS, AGS\(l\) & CRS & DE\\
+    & AGS & CRS & DE\\
   \hline
@@
 \begin{center}
-\caption{Number of test optimization problems solved by methods}
+\caption{Number of test optimization problems solved by the methods}
   \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
@@
   \item the proposed modification of the stock AGS, AGS-AR allows to pay less attention to initial hyperparameter tuning and
-  performs on-par with properly tuned AGS.
+  performs on-par with properly tuned AGS;
   \item AGS-AR method has demonstrated the convergence
@@
 speed and in reliability. It is manifested especially strongly on more complex multiextremal
-problems;
+problems.
 \end{itemize}

### 2019-02-28 2019/NUMTA_Barkalov/Abstract/Barkalov.tex

% Latex-template for one-page abstract submissions.
% NUMTA2019 - Numerical Computations: Theory and Algorithms
% 15-21 June 2019, Le Castella - Isola Capo Rizzuto, Crotone (Italy)

\documentclass[oribibl]{llncs}

% Please, do not use your own macros/redefinitions.

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}
\cleardoublepage

\title{Title of Your Presentation}
\author{Roman Strongin, Konstantin Barkalov, Semen Bevzuk}

%%% Please note that the first author is supposed to be the speaker.

\institute{University of Nizhni Novgorod, Nizhni Novgorod, Russia\\
\email{president@unn.ru, barkalov@vmk.unn.ru}}

\maketitle

\thispagestyle{fancy}

\textbf{Keywords.} Global optimization; keyword2; keyword3.

 \vspace*{0.5cm}

%-----------------------------------------------------------------

The goal of the Conference is to create a multidisciplinary round
table for an open discussion on numerical modeling nature by using
traditional and emerging computational paradigms. The Conference
will discuss all aspects of numerical computations and modeling from
foundations and philosophy to advanced numerical techniques. New
technological challenges and fundamental ideas from theoretical
computer science, linguistic, logic, set theory, and philosophy will
meet requirements and new fresh applications from physics,
chemistry, biology, and economy.

Researchers from both theoretical and applied sciences are very
welcome to submit their papers and to use this excellent possibility
to exchange ideas with leading scientists from different research
fields. Papers discussing new computational paradigms, relations
with foundations of mathematics, and their impact on natural
sciences are particularly solicited. A special attention will be
also dedicated to numerical optimization and different issues
related to theory and practice of the usage of infinities and
infinitesimals in numerical computations.

\textbf{Acknowledgements.}

This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.

\vspace{0.5cm}

\begin{thebibliography}{4}

\bibitem{Book} Strongin R.G., Sergeyev Ya.D. (2000) \emph{Global optimization with non-convex constraints. Sequential and parallel algorithms}. Kluwer Academic Publishers, Dordrecht.

\bibitem{Strongin18} Barkalov K., Strongin R. (2018) Solving a set of global optimization problems by the parallel technique with uniform convergence. \emph{J. Glob. Optim.}, Vol.~71(1), pp.~21--36.

\bibitem{Sergeyev16} Sergeyev Ya.D., Mukhametzhanov M.S., Kvasov D.E., Lera D. (2016) Derivative-free local tuning and local improvement techniques embedded in the univariate global optimization. \emph{J. Optim. Theory Appl.}, Vol.~171(1), pp.~186--208.

\bibitem{Sergeyev10} Lera D., Sergeyev Ya.D. (2010)  (2010) An information global minimization algorithm using the local improvement technique. \emph{J. Glob. Optim.}, Vol.~48(1), pp.~99--112.

\end{thebibliography}

\end{document}

### 2019-03-06 2019/NUMTA_Barkalov/Abstract/Barkalov.tex
Russian text

@@
 
+\usepackage[utf8x]{inputenc}
+\usepackage[english,russian]{babel}
+\usepackage{cmap}
+
 \begin{document}
@@
 
-\title{Title of Your Presentation}
+\title{A new way to speedup in global optimization problems}
 \author{Roman Strongin, Konstantin Barkalov, Semen Bevzuk}
@@
 \institute{University of Nizhni Novgorod, Nizhni Novgorod, Russia\\
-\email{president@unn.ru, barkalov@vmk.unn.ru}}
+\email{barkalov@vmk.unn.ru}}
 
@@
 
-\textbf{Keywords.} Global optimization; keyword2; keyword3.
+\textbf{Keywords.} Global optimization; multiextremal problems; Lipschitz constant.
 
@@
 
-The goal of the Conference is to create a multidisciplinary round
-table for an open discussion on numerical modeling nature by using
-traditional and emerging computational paradigms. The Conference
-will discuss all aspects of numerical computations and modeling from
-foundations and philosophy to advanced numerical techniques. New
-technological challenges and fundamental ideas from theoretical
-computer science, linguistic, logic, set theory, and philosophy will
-meet requirements and new fresh applications from physics,
-chemistry, biology, and economy.
+В работе рассматриваются задачи глобальной оптимизации c black-box objective function, удовлетворяющей условию Липшица.
+Важной проблемой в задачах данного класса является адекватная оценка априори неизвестной константы Липшица на основе результатов проведенных search trials.
+В монографии \cite{Book} предложены различные подходы к учету как глобальных, так и локальных свойств objective function.
+В частности, показали свою перспективность  алгоритмы с использованием локальных оценок константы Липшица \cite{Sergeyev10, Sergeyev16}.
 
-Researchers from both theoretical and applied sciences are very
-welcome to submit their papers and to use this excellent possibility
-to exchange ideas with leading scientists from different research
-fields. Papers discussing new computational paradigms, relations
-with foundations of mathematics, and their impact on natural
-sciences are particularly solicited. A special attention will be
-also dedicated to numerical optimization and different issues
-related to theory and practice of the usage of infinities and
-infinitesimals in numerical computations.
+В данной статье предлагается новый способ учета локальных свойств objective function.
+Предложенный способ основан на использовании в информационно-статистическом алгоритме глобального поиска \cite{Book} одновременно двух оценок константы Липшица: нижней и верхней.
+Правила нового алгоритма обеспечивают автоматическое переключение между указанными двумя оценками в зависимости от поведения оптимизируемой функции и от стадии поиска.
+
+Предложенный новый алгоритм гарантирует сходимость к global minimizer.
+Результаты решения серии тестовых задач разной размерности подтверждают как наличие глобальной сходимости  нового алгоритма, так и его ускорение по сравнению с исходным алгоритмом.
 
@@
 
-\bibitem{Strongin18} Barkalov K., Strongin R. (2018) Solving a set of global optimization problems by the parallel technique with uniform convergence. \emph{J. Glob. Optim.}, Vol.~71(1), pp.~21--36.
+\bibitem{Sergeyev10} Lera D., Sergeyev Ya.D. (2010)  (2010) An information global minimization algorithm using the local improvement technique. \emph{J. Glob. Optim.}, Vol.~48(1), pp.~99--112.
 
@@
 
-\bibitem{Sergeyev10} Lera D., Sergeyev Ya.D. (2010)  (2010) An information global minimization algorithm using the local improvement technique. \emph{J. Glob. Optim.}, Vol.~48(1), pp.~99--112.
+%\bibitem{Strongin18} Barkalov K., Strongin R. (2018) Solving a set of global optimization problems by the parallel technique with uniform convergence. \emph{J. Glob. Optim.}, Vol.~71(1), pp.~21--36.
 

### 2019-03-07 2019/NUMTA_Barkalov/Abstract/Barkalov.tex
First English version

@@
 
-\usepackage[utf8x]{inputenc}
-\usepackage[english,russian]{babel}
-\usepackage{cmap}
+%\usepackage[utf8x]{inputenc}
+%\usepackage[english,russian]{babel}
+%\usepackage{cmap}
 
@@
 
-В работе рассматриваются задачи глобальной оптимизации c black-box objective function, удовлетворяющей условию Липшица.
-Важной проблемой в задачах данного класса является адекватная оценка априори неизвестной константы Липшица на основе результатов проведенных search trials.
-В монографии \cite{Book} предложены различные подходы к учету как глобальных, так и локальных свойств objective function.
-В частности, показали свою перспективность  алгоритмы с использованием локальных оценок константы Липшица \cite{Sergeyev10, Sergeyev16}.
+The paper considers global optimization problems with a black-box objective function satisfying a Lipschitz condition. 
+An important issue in this class of problems is reliable estimates of a priori unknown Lipschitz constant based on the results of the search trials. 
+In fundamental work \cite{Book} various approaches have been proposed to take into account both global and local properties of the objective function. 
+%В частности, показали свою перспективность  алгоритмы с использованием локальных оценок константы Липшица 
+In particular, algorithms using local estimates of the Lipschitz constant have shown their promise \cite{Sergeyev10, Sergeyev16}.
 
-В данной статье предлагается новый способ учета локальных свойств objective function.
-Предложенный способ основан на использовании в информационно-статистическом алгоритме глобального поиска \cite{Book} одновременно двух оценок константы Липшица: нижней и верхней.
-Правила нового алгоритма обеспечивают автоматическое переключение между указанными двумя оценками в зависимости от поведения оптимизируемой функции и от стадии поиска.
+This paper proposes a new way to account for the local properties of the objective function.
+The proposed method is based on the use in the information-statistical global search algorithm \cite{Book} lower and upper estimates of the Lipschitz constant simultaneously. 
+The rules of the new algorithm provide automatic switching between these two estimates, depending on the type of the objective function and on the search stage.
 
-Предложенный новый алгоритм гарантирует сходимость к global minimizer.
-Результаты решения серии тестовых задач разной размерности подтверждают как наличие глобальной сходимости  нового алгоритма, так и его ускорение по сравнению с исходным алгоритмом.
+Proposed algorithm guarantees convergence to global minimizer.
+The results of solving several series of multiextremal test problems confirm both the global convergence of the new algorithm and its acceleration compared to the original one.
 

### 2019-03-07 2019/NUMTA_Barkalov/Abstract/Barkalov.tex
Second English version

@@
 
-\title{A new way to speedup in global optimization problems}
+\title{Acceleration of global search through dual Lipschitz constant estimates}
 \author{Roman Strongin, Konstantin Barkalov, Semen Bevzuk}
@@
 
-\textbf{Keywords.} Global optimization; multiextremal problems; Lipschitz constant.
+\textbf{Keywords.} Global optimization; multiextremal problems; Lipschitz constant estimates.
 
@@
 
-The paper considers global optimization problems with a black-box objective function satisfying a Lipschitz condition. 
-An important issue in this class of problems is reliable estimates of a priori unknown Lipschitz constant based on the results of the search trials. 
+The paper considers global optimization problems with a black-box objective function satisfying the Lipschitz condition. 
+Efficient algorithms for this class of problems require reliable estimates of the Lipschitz constant to be introduced. 
 In fundamental work \cite{Book} various approaches have been proposed to take into account both global and local properties of the objective function. 
 %В частности, показали свою перспективность  алгоритмы с использованием локальных оценок константы Липшица 
-In particular, algorithms using local estimates of the Lipschitz constant have shown their promise \cite{Sergeyev10, Sergeyev16}.
+In particular, algorithms using local estimates of the Lipschitz constant have shown their potential \cite{Sergeyev10, Sergeyev16}.
 
-This paper proposes a new way to account for the local properties of the objective function.
-The proposed method is based on the use in the information-statistical global search algorithm \cite{Book} lower and upper estimates of the Lipschitz constant simultaneously. 
-The rules of the new algorithm provide automatic switching between these two estimates, depending on the type of the objective function and on the search stage.
-
-Proposed algorithm guarantees convergence to global minimizer.
-The results of solving several series of multiextremal test problems confirm both the global convergence of the new algorithm and its acceleration compared to the original one.
+The new approach presented in this paper is based on simultaneous use of two estimates: one is substantially larger then the other. 
+The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
+Results of numerical experiments on the random sample of multidimensional functions demonstrate the efficiency of the suggested approach.
 

### 2019-03-18 2019/PAVT/slides/slides.tex
PAVT: add a placeholder for slides

--- 2018-09-24 2018/RSD/MultiEvolvents/slides/slides.tex
@@
 
-\graphicspath{ {../pictures/}{img/} }
+\graphicspath{ {../images/}{img/} }
 
@@
 
-\title{Comparison of Dimensionality Reduction Schemes for Parallel Global Optimization Algorithms}
-\author{\textbf{Konstantin~Barkalov}, \underline{\textbf{Vladislav~Sovrasov}} and \textbf{Ilya~Lebedev}}
+\title{Comparison Of Several Sequential And Parallel Derivative-free Global Optimization Algorithms}
+\author{\underline{\textbf{Vladislav~Sovrasov}} and \textbf{Semen~Bevzuk}}
 \institute{Lobachevsky State University of Nizhni Novgorod}
@@
 \begin{frame}
-  \frametitle{Non-univalent evolvent}
-    One can try to recover all preimages of \(y\in\mathbb{R}^N\) and make optimization method aware of their existence\footnote{R.G.
-Strongin.
-Numerical Methods in Multiextremal Problems (in Russian), 1978}.
-This allows reducing the effect of growing amount of local minimas after dimension reduction.
-    According to the theory of Peano-type curves, each \(N\)-d point could have up to \(2^N\) preimages. For large \(N\) such preimages mining would be expensive.
-      \begin{figure}[ht]
-        \includegraphics[width=.25\textwidth]{noninjective.pdf}
-      \end{figure}
-\end{frame}
+  \frametitle{Basic optimization method}
 
-\begin{frame}
-  \frametitle{Shifted and rotated evolvents}
-  To create a fixed amount of preimages one can use a pre-defined set of different evolvents.
-These evolvents could be shifted or rotated versions of the original one.
-Set of shifted evolvents\footnote{Strongin, R.G.
-Algorithms for multi-extremal mathematical programming problems employing the set of joint space-filling curves, Journal of Global Optimization 2(4), 357--378, 1992} is theoretically proven to generate at least one pair of close preimages if their images are close and it performs better than the a set of rotated curves in that sense.
-  \begin{figure}[ht]
-    \subfloat{\raisebox{.0\textheight}{\includegraphics[width=.35\textwidth]{shifted.pdf} }}
-    \subfloat{{\includegraphics[width=.35\textwidth]{rotated.pdf} }}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Smooth evolvent}
-  Smooth functions are more predictable for optimizer, so smooth approximation of the Peano-like \(y(x)\) curve could improve convergence rate \footnote{Goryachih, A.
-A class of smooth modification of space-filling curves for global optimization problems, NET 2016}.
-  \begin{figure}[ht]
-    \includegraphics[width=0.45\textwidth]{smooth.pdf}
-  \end{figure}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Basic parallel optimization method}
   Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
@@
   \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Parallel optimization method with multiple evolvents}
-  Using the multiple mapping allows solving initial problem by parallel solving the problems
-  \[
-  \min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
-  \]
-  on a set of intervals $[0,1]$ by the basic method. Each one-dimensional problem is solved on a
-  separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
-  particular processor are interpreted as the results of the trials in the rest problems (in the
-  corresponding points \(x^{k_1},\dots,x^{k_S})\). In this approach, a trial at the point \(x^k \in
-  [0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
-  operations:
-  \begin{enumerate}
-    \setlength{\itemindent}{.1in}
-    \item[Step 1.] Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
-    \item[Step 2.] Inform the rest of processors about the start of the trial execution at the point \( y^k\) (the
-    blocking of the point \(y^k\) ).
-    \item[Step 3.] Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
-    trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
-    \(x{}^{k_1} ,\dots,x{}^{k_s} \)
-    \item[Step 4.] Inform the rest of processors about the trial results at the point \(y^k\).
-  \end{enumerate}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Parallel optimization method with multiple evolvents}
-  \begin{figure}[ht]
-    \includegraphics[width=0.95\textwidth]{evolvents_parallel.pdf}
-  \end{figure}
 \end{frame}
@@
   \frametitle{Evolvents comparison}
-  \begin{figure}[ht]
-    \hspace*{-0.9cm}
-    \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_opt_pt_op.pdf} }}
-    \subfloat[\(r=5.0\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_same_r_opt_pt_op.pdf} }} \hspace*{4cm}
-    \caption{Operating characteristics on GKLS 2d Simple class}
-  \end{figure}
+  %\begin{figure}[ht]
+  %  \hspace*{-0.9cm}
+  %  \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_opt_pt_op.pdf} }}
+  %  \subfloat[\(r=5.0\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_same_r_opt_pt_op.pdf} }} \hspace*{4cm}
+  %  \caption{Operating characteristics on GKLS 2d Simple class}
+  %\end{figure}
 \end{frame}
@@
   \frametitle{Evolvents comparison}
-  \begin{figure}[ht]
-    \hspace*{-0.9cm}
-    \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_opt_pt_op.pdf} }}
-    \subfloat[\(r=4.5\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_same_r_opt_pt_op.pdf} }}
-    \caption{Operating characteristics on GKLS 3d Simple class}
-  \end{figure}
+  %\begin{figure}[ht]
+  %  \hspace*{-0.9cm}
+  %  \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_opt_pt_op.pdf} }}
+  %  \subfloat[\(r=4.5\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_same_r_opt_pt_op.pdf} }}
+  %  \caption{Operating characteristics on GKLS 3d Simple class}
+  %\end{figure}
 \end{frame}
@@
   \frametitle{Choice of evolvent for the parallel algorithm}
-  \begin{itemize}
-    \item Smooth evolvent is too computational heavy.
-    \item Non-univalent evolvent generates large and unpredictable amount of preimages.
-    \item Shifted evolvent generates huge amount of auxiliary points to handle additional constraint.
-  \end{itemize}
-  \begin{table}
-  \begin{center}
-  \caption{Averaged number of computations of \(g_0\) and of \(\varphi\) when solving the
-  problems from GKLS 3d Simple class using the shifted evolvent}
-    \begin{tabular}{|l|{c}|{c}|{c}|}
-      \hline
-    $L$ & $calc(g_0)$ & $calc(\varphi)$ & $\frac{calc(g_0)}{calc(\varphi)}$ ratio \\
-    \hline
-    2 & 96247.9  & 6840.14 & 14.07\\
-    \hline
-    3 & 153131.0 & 7702.82 & 19.88\\
-    \hline
-    \end{tabular}
-    \label{tab:shifted_g0}
-  \end{center}
-  \end{table}
 \end{frame}
@@
 
-  \begin{table}
-    \centering
-    \caption{Averaged numbers of iterations executed by the parallel algorithm for solving the test
-  optimization problems}
-    \label{tab:iterations}
-    \begin{tabular}{cccccccc}
-      \cline{3-8}\noalign{\smallskip}
-      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & &
-  \multicolumn{2}{c}{$N=5$}   \\
-      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
-      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
-  \textit{Hard}  \\
-      \noalign{\smallskip}\hline
-      I &
-      \parbox{0.25\textwidth}{
-      \begin{center}
-      \textbf{1 cluster node}
-      \end{center}		}
-        & \textit{1} & 12167 & 25635 & & 20979 & 187353  \\
-      &  & \textit{32} & 328 & 1268  & &   898 & 12208 \\
-      \hline \noalign{\smallskip}
-  II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
-    & \textit{1} & 25312 & 11103 & & 1472 & 17009 \\
-  &   & \textit{32} & 64 &   913 & & 47 & 345 \\
-      \noalign{\smallskip}\hline	\noalign{\smallskip}
-  III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-    & \textit{1}  & 810 & 4351 & & 868 & 5697  \\
-  & & \textit{32} & 34  & 112  & & 35  & 868 \\
-      \noalign{\smallskip}\hline
-    \end{tabular}
-  \end{table}
-
 \end{frame}
@@
   \frametitle{Results of applying the parallel algorithm}
-
-  \begin{table}
-    \centering
-    \caption{Speedup of parallel computations executed by the parallel algorithm}
-    \label{tab:speedup}
-    \begin{tabular}{cccccccc}
-      \cline{3-8}\noalign{\smallskip}
-      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & &
-  \multicolumn{2}{c}{$N=5$}   \\
-      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
-      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
-  \textit{Hard}  \\
-      \noalign{\smallskip}\hline
-      I &
-      \parbox{0.25\textwidth}{
-      \begin{center}
-      \textbf{1 cluster node}
-      \end{center}		}
-      & \textit{1}   & 12167(10.58s) & 25635(22.26s) & & 20979(22.78s) & 187353(205.83s)  \\
-    &  & \textit{32} & 37.1(18.03) & 20.2(8.55)  & &  23.3(8.77) & 15.4(9.68) \\
-    \hline \noalign{\smallskip}
-  II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
-  & \textit{1} &        0.5(0.33) & 2.3(0.86)  & & 14.3(6.61) & 11.0(6.06) \\
-  &   & \textit{32} & 190.1(9.59) & 28.1(1.08) & & 446.4(19.79) & 543.0(43.60) \\
-    \noalign{\smallskip}\hline	\noalign{\smallskip}
-  III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-  & \textit{1}    & 15.0(6.05)  & 5.9(2.36)   & & 24.2(17.56)  & 32.9(24.87)  \\
-  & & \textit{32} & 357.9(2.36) & 228.9(2.64) & & 582.8(20.96) & 793.0(33.89) \\
-      \noalign{\smallskip}\hline
-    \end{tabular}
-  \end{table}
 
@@
     \begin{itemize}
-      \item The smooth evolvent and the non-univalent one demonstrate the best result in the problems of small dimensionality and can be applied successfully in solving the problems with the computational costly objective functions.
-      \item The shifted evolvents introduce large overhead costs on the execution of the method due to the requirement to adding an auxiliary constraint.
-About 95\% of iterations are overhead to fight the auxiliary constraint.
-      \item Rotated evolvents perform almost the same as the shifted ones but without overhead.
-      \item Parallel optimization method shows up to 43x speedup on hard \(5d\) problems when using a set of rotated evolvents.
+      \item 1
     \end{itemize}

### 2019-03-19 2019/WCGO/WCGO.tex
Final

@@
 \end{equation}
-unambiguously mapping the interval of real axis $[0,1]$ onto a 
+unambiguously mapping the interval of real axis $[0,1]$ onto an 
 $N$-dimensional cube is the first of the dimension reduction methods considered.
@@
 numbers of trials executed by GSA with the use of evolvents ($K_e$), nested optimization scheme ($K_n$), and adaptive 
-nested optimization scheme ($K_a$). Fig.~\ref{fig1} and Fig.~\ref{fig2}(a,b) present the operating characteristics of the 
+nested optimization scheme ($K_{an}$). Fig.~\ref{fig1} and Fig.~\ref{fig2}(a,b) present the operating characteristics of the 
 algorithms obtained on the problem classes $F_{gr}$, GKLS \textit{Simple}, and GKLS \textit{Hard} respectively. The solid 
-line corresponds to the algorithm using the evolvents, short dashed line -- to the adaptive nested optimization scheme, and 
-the long dashed line -- to the nested optimization scheme. The results of experiments demonstrate that GSA with the use of 
+line corresponds to the algorithm using the evolvents (GSA-E), short dashed line -- to the adaptive nested optimization scheme (GSA-AN), and 
+the long dashed line -- to the nested optimization scheme (GSA-N). The results of experiments demonstrate that GSA with the use of 
 the adaptive nested optimization scheme shows almost the same speed as compared to GSA with the evolvents, and 
@@
  $K_n$ & 341  & 697 & 1252 \\
- $K_a$ & 215  & 279 & 815 \\
+ $K_{an}$ & 215  & 279 & 815 \\
 \noalign{\smallskip}\hline
@@
 \textit{Simple} and GKLS \textit{Hard} (100 functions of each class). Table~\ref{tab2}  presents the averaged numbers 
-of trials executed by GSA with the use of the adaptive nested optimization scheme ($K_a$) and block adaptive nested 
-optimization scheme ($K_{ba}$) with two levels of subproblems of equal dimensionality $N_1=N_2=2$. 
+of trials executed by GSA with the use of the adaptive nested optimization scheme ($K_{an}$) and block adaptive nested 
+optimization scheme ($K_{ban}$) with two levels of subproblems of equal dimensionality $N_1=N_2=2$. 
 Note that when solving the problem of the dimensionality $N=4$ using the initial variant of the adaptive scheme, four 
@@
 and GKLS \textit{Hard} respectively. The dashed line corresponds to GSA using the adaptive nested optimization 
-scheme, the solid line -- the block adaptive nested optimization scheme. 
+scheme (GSA-AN), the solid line -- the block adaptive nested optimization scheme (GSA-BAN). 
 The results of experiments demonstrate the use of the block adaptive nested optimization scheme provides a considerable 
@@
 \noalign{\smallskip}\hline\noalign{\smallskip}
- $K_a$  &  21747 & 35633 \\
- $K_{ba}$ &  13894 & 31620 \\
+ $K_{an}$  &  21747 & 35633 \\
+ $K_{ban}$ &  13894 & 31620 \\
 \noalign{\smallskip}\hline

### 2019-03-19 2019/WCGO/WCGO.tex
Final

@@
 %
-\title{Adaptive global optimization based on nested dimensionality 
-reduction\thanks{This study was supported by the Russian Science Foundation, 
+\title{Adaptive Global Bptimization Based on Nested Dimensionality 
+Reduction\thanks{This study was supported by the Russian Science Foundation, 
 project No.\,16-11-10150.}}
 %
-\titlerunning{Adaptive global optimization}
+\titlerunning{Adaptive Global Optimization}
 % If the paper title is too long for the running head, you can set 

### 2019-03-19 2019/WCGO/WCGO.tex
Final!

@@
 %
-\title{Adaptive Global Bptimization Based on Nested Dimensionality 
+\title{Adaptive Global Optimization Based on Nested Dimensionality 
 Reduction\thanks{This study was supported by the Russian Science Foundation, 

### 2019-03-25 2019/RSD_Integer/RSD2019.tex

--- 2018-04-19 2018/LION_Mixed-Integer/LION2018.tex
@@
 
-\title{A Global Optimization Algorithm for Non-Convex Mixed-Integer Problems}
-\author{Victor Gergel \and Konstantin Barkalov \and Ilya Lebedev %\Letter 
+\title{Parallel Global Optimization for Non-Convex Mixed-Integer Problems}
+\author{Konstantin Barkalov \and Ilya Lebedev %\Letter 
 \\
@@
 
-\section{Global search algorithm and dimension reduction}
+\section{Global optimization algorithm and dimension reduction}
 
@@
 \begin{gather}\label{problem}
-\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
+\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_j(y)\leq 0, \; 1 \leq j \leq m\right\}},\\
 D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.\nonumber
@@
 Suppose, that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and
-the left-hand sides $g_i(y), \; 1\leq i \leq m$, of the constraints satisfy the Lipschitz condition
-%\[ \left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq i \leq m+1, \]
-%with constants $L_i, \; 1 \leq i \leq m+1$, respectively,  
+the left-hand sides $g_j(y), \; 1\leq j \leq m$, of the constraints satisfy the Lipschitz condition
+\[ \left|g_j(y')-g_j(y'')\right| \leq L_j \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq j \leq m+1, \]
+with constants $L_j, \; 1 \leq j \leq m+1$, respectively,  
 and may be multiextremal.
@@
 \begin{equation}\label{problem1}
-%\varphi(y(x^\ast))=
-\min \left\{g_{m+1}(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
+\varphi(y(x^\ast))= \min \left\{g_{m+1}(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
 \end{equation}
@@
 An efficient global search algorithm (GSA) for solving the constrained optimization problem 
-(\ref{problem1}) has been developed at University of Nizhni Novgorod. A detailed description 
+(\ref{problem1}) has been developed at University of Nizhni Novgorod.
+
+%Описание индексного алгоритма
+
+A detailed description 
 of the algorithm and the corresponding convergence theory are presented in 
@@
 \begin{gather}\label{problem_i}
-\min{\left\{ g_{m+1}(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
+\min{\left\{ g_{m+1}(y):y\in D, \; g_j(y)\leq 0, \; 1 \leq j \leq m\right\}},\\
 D=\left\{a_i\leq y_i \leq b_i, \; 1\leq i \leq N, \; y_j \in Z, \; j \in J, \; y_i \in R, \; i \notin J \right\}.\nonumber

### 2019-03-25 2019/RSD_Integer/RSD2019.tex

@@
 %\usepackage{epsfig}
+
+\usepackage[utf8x]{inputenc}
+\usepackage[english,russian]{babel}
 
@@
 
-Now, let us consider a method of adaptation of GSA for solving the mixed-integer global 
-optimization problems 
+Рассмотрим теперь случай, когда аргумент функций задачи содержит две компоненты: вектор $y$, принадлежащий гиперинтервалу $D$, и вектор $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е.
 \begin{gather}\label{problem_i}
-\min{\left\{ g_{m+1}(y):y\in D, \; g_j(y)\leq 0, \; 1 \leq j \leq m\right\}},\\
-D=\left\{a_i\leq y_i \leq b_i, \; 1\leq i \leq N, \; y_j \in Z, \; j \in J, \; y_i \in R, \; i \notin J \right\}.\nonumber
+\min{\left\{ g_{m+1}(y,u):y\in D, \; g_j(y,u)\leq 0, \; 1 \leq j \leq m\right\}},\\
+D=\left\{a_i\leq y_i \leq b_i, \; 1\leq i \leq N \right\}.\nonumber
 \end{gather}
+
+Такие конечные наборы могут характеризовать, например, вариант материала, из которого создается объект, геометрические размеры или другие величины, которые могут принадлежать стандартному дискретному ряду, и т.п.
+
+Занумеруем целыми числами $s, 1\leq s \leq S,$ все возможные значения вектора $u$, т.е. сопоставим каждому рассматриваемому значению $s$ вектор $u_s$. 
+Тогда рассматриваемая задача может быть записана в виде 
+\begin{gather}\label{problem_is}
+ \min_{s\in\{1,...,S\}}\left(\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_j(y,u_s)\leq 0, \; 1 \leq j \leq m\right\}}\right),\\
+D=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, 1 \leq i\leq N \right\}.\nonumber 
+\end{gather}
+
+Используя схему редукции размерности с помощью развертки $y(x), x\in [0,1]$ можно сопоставить каждой вложенной задаче минимизации по $y$ одномерную задачу минимизации
 

### 2019-03-25 2019/RSD_Integer/RSD2019.tex

@@
 Используя схему редукции размерности с помощью развертки $y(x), x\in [0,1]$ можно сопоставить каждой вложенной задаче минимизации по $y$ одномерную задачу минимизации
+\[
+ \min{\left\{ g_{m+1}(y(x),u_s):x \in [0,1], \; g_j(y(x),u_s)\leq 0, \; 1 \leq j \leq m\right\}}, s\in\{1,...,S\}.
+\]
 
-Let us formulate a set of problems based on the problem (\ref{problem_i})
-\begin{gather}\label{problem_is}
-\min{\left\{ g_{m+1}^s(y):y\in D, \; g_i^s(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
-D=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, \; i \notin J \right\}, \; s\in\{1,...,S\}.\nonumber 
-\end{gather}
-Each of these problems corresponds to the initial problem (\ref{problem_i}) with some combination of 
-integer parameters. The number of problems $S$ will correspond to the number of all possible 
-combinations of the integer parameters.
-
-Using the dimensionality reduction scheme utilizing the evolvent $y(x)$ 
-%based on the set of problems (\ref{problem_is}),
-one can compose a united problem 
-\begin{equation}\label{problem_is1}
-\min \left\{g_{m+1}^s(y^s(x)): x \in [0,S], \; g_i^s(y^s(x)) \leq 0, \; 1 \leq i \leq m\right\},
-\end{equation}
-where the mapping $y^s(x)$ is formed on the basis of initial evolvent $y(x)$ as follows
+Рассмотрим теперь соответствие 
 \[
-y^s(x)=y(x-s), \; x\in[s-1,s],\; s\in\{1,...,S\}.
+Y(x)=y(x-E(x)), \; x\in[0,S],
 \]
-The functions $g_i^s(y^s(x))$ can have the breaks at the points $s\in \{1,...,S-1\}$, therefore, 
-these points were considered as the punctured ones, the values of the objective function and 
-constraints at these points remain undefined.
+отображающее любую точку интервала [0,S] на область $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функции 
+\[
+g_j(x) = g_j(Y(x),u_{E(x)+1}), x\in[0,S],
+\]
+имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_i = i, 1\leq i \leq S-1$.
+Используя введенные обозначения можно переформулировать исходную задачу как
+\[
+\min \left\{g_{m+1}(x): x \in [0,S], \; g_j(x) \leq 0, \; 1 \leq j \leq m\right\}.
+\]
 

### 2019-03-27 2019/RSD_Integer/RSD2019.tex

@@
 
-Let us consider a multiextremal optimization problem in the form
+A constrained global optimization problem can be formulated as follows
+%Let us consider the $N$-dimensional global optimization problem
 \begin{gather}\label{problem}
-\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_j(y)\leq 0, \; 1 \leq j \leq m\right\}},\\
-D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.\nonumber
+\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
+D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.\label{D}
 \end{gather}
-Suppose, that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and
-the left-hand sides $g_j(y), \; 1\leq j \leq m$, of the constraints satisfy the Lipschitz condition
-\[ \left|g_j(y')-g_j(y'')\right| \leq L_j \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq j \leq m+1, \]
-with constants $L_j, \; 1 \leq j \leq m+1$, respectively,  
-and may be multiextremal.
+The objective function $\varphi(y)$ (hereafter denoted by $g_{m+1}(y)$) and the left-hand 
+sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
+\[
+\left|g_i(y_1)-g_i(y_2)\right|\leq L_i\left\|y_1-y_2\right\|, \;1\leq i\leq m+1, \; y_1,y_2 \in D,\;
+\]
+with a priori unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal. It is 
+assumed that the functions $g_i(y)$ are defined and computable only at the points $y \in D$ 
+satisfying the conditions
+\begin{equation}\label{g_k}
+g_k(y) \leq 0, \; 1 \leq k < i.
+\end{equation}
 
-Using a continuous single-valued mapping $y(x)$  (Peano-type space-filling curve, or \textit{evolvent}) of the interval $[0,1]$ onto $D$, a multidimensional problem (\ref{problem}) can be reduced to a one-dimensional problem
-\begin{equation}\label{problem1}
-\varphi(y(x^\ast))= \min \left\{g_{m+1}(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
+By employing the continuous single-valued Peano curve $y(x)$ mapping the unit interval 
+$[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}), it is possible to find the 
+minimum in (\ref{problem}) by solving the one-dimensional problem
+\[
+\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
+\]
+Algorithms for numerical construction of Peano curve approximation (\textit{evolvent}) are 
+considered in \cite{Strongin2013}. Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined 
+and computable in the subranges 
+\[
+Q_1=[0,1], \; Q_{i+1}=\left\{x \in Q_i : g_i(y(x)) \leq 0 \right\}, \; 1 \leq i \leq m.
+\]
+
+These conditions allows us to introduce a classification of the points $x \in [0,1]$ according to 
+the number $\nu (x)$ of the constraints computed at this point. The \textit{index} $\nu(x)$ can 
+also be defined by the conditions
+\begin{equation}\label{nu}
+g_i(y(x)) \leq 0, \; 1 \leq i < \nu, \; g_\nu(y(x))>0,
 \end{equation}
-The reduction of dimensionality matches the multidimensional problem with a Lipschitzian objective function and constraints with a one-dimensional problem where the respective functions satisfy the uniform H\"older condition (see \cite{Strongin2000}).
+where the last inequality is inessential if $\nu=m+1$.
 
-An efficient global search algorithm (GSA) for solving the constrained optimization problem 
-(\ref{problem1}) has been developed at University of Nizhni Novgorod.
+In the dimensionality reduction scheme considered here, a multidimensional problem with 
+Lipschitzian functions is juxtaposed with a one-dimensional problem, where the corresponding 
+functions satisfy uniform H{\"o}lder condition (see \cite{Strongin2013}), i.e.,
+\[
+\left|g_i(y(x_1))-g_i (y(x_2))\right| \leq H_i \left|x_1-x_2 \right|^{1/N}, \; x_1,x_2\in [0,1], \; 
+1\leq i \leq m+1.
+\]
+Here, $N$ is the dimensionality of the initial multidimensional problem and the coefficients 
+$H_i$ are related to the Lipschitz constant $L_i$ of the initial problem as $H_i \leq 2L_i 
+\sqrt{N+3}$.
 
-%Описание индексного алгоритма
+Thus, \textit{a trial} at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm 
+will consist of the following sequence of operations:
+\begin{itemize}
+	\item Determine the \textit{image} $y^k=y(x^k)$ in accordance with the mapping 
+$y(x)$;
+	\item Compute the values $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from 
+(\ref{nu}). 
+\end{itemize}
+The dyad $ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} $ will be referred to as the \textit{trial 
+outcome}.
 
-A detailed description 
-of the algorithm and the corresponding convergence theory are presented in 
-\cite{Strongin2000}.
+An efficient parallel index algorithm (PIA) for solving the constrained optimization problem 
+(\ref{problem}) has been developed at University of Nizhni Novgorod.
 
-\section{Global search algorithm for mixed-integer problems}
+Suppose we have $p \geq 1$  computational elements (e.g., accelerator cores), which can be used to run $p$ trials simultaneously.
+In the first iteration of the method, $p$ trials are run in parallel at various random points $x^i\in(0,1)$, $1\leq i \leq p$.
+Suppose $n \geq 1$  iterations of the method have been completed, and as a result of which, trials were carried out in $k=k(n)$ points $x^i, 1\leq i \leq k$.
+Then the points $x^{k+1},...,x^{k+p}$  of the search trials in the next $(n+1)$-th iteration will be determined according to the rules below.
+
+\begin{enumerate}
+\item 
+Renumber the points $x^1,...,x^k$ from previous iterations with lower indices, lowest to highest coordinate values, i.e.
+\begin{equation}\label{Eq:17}
+0=x_0<x_1<...<x_i<...<x_k<x_{k+1}=1,
+\end{equation}
+and match them with the values $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, from (\ref{Eq:16}) , calculated at these points; points $x_0=0$ и $x_{k+1}=1$ are introduced additionally; the values $z_0$ и $z_{k+1}$ are indeterminate.
+\item
+Classify the numbers $i,1\leq i \leq k$, of the trial points from (\ref{Eq:17}) by the number of problem constraints fulfilled at these points, by building the sets
+\begin{equation}\label{Eq:18}
+I_\nu = \left\{i: 1 \leq i \leq k,\ \nu = \nu(x_i)\right\},\ 1 \leq \nu \leq m+1,
+\end{equation}
+containing the numbers of all points $x_i,1\leq i \leq k$, with the same values of $\nu$.
+The end points $x_0=0$ and $x_{k+1}=1$ are interpreted as those with zero indices, and they are matched to an additional set $I_0={0,k+1}$.
+
+Identify the maximum current value of the index
+\begin{equation}\label{Eq:19}
+M=\max \left\{\nu = \nu(x_i), \ 1\leq i \leq k\right\}.
+\end{equation}
+\item
+For all values of $\nu, \ 1\leq \nu \leq m+1$, calculate the values  
+\begin{equation}\label{Eq:20}
+\mu_\nu = \max \left\{ \frac{\left|z_i-z_j\right|}{\left(x_i-x_j\right)^{1/N}} : i,j \in I_\nu, j<i\right\}.
+\end{equation}
+If the set $I_\nu$ contains less than two elements or $\mu_\nu$ from (\ref{Eq:20}) equals zero, then assume $\mu_\nu=1$.
+\item
+For all non-empty sets $I_\nu$, $1 \leq \nu \leq m+1$, determine the values
+\begin{equation}\label{Eq:21}
+  z^\ast_\nu =  
+   \begin{cases}
+    -\epsilon_\nu,  \nu < M, \\
+    \min{\left\{g_\nu(x_i):i\in I_\nu\right\}}, \nu = M,
+   \end{cases}
+\end{equation}
+where $M$ is the maximum current value of the index, and the vector
+\begin{equation}\label{Eq:22}
+\epsilon _R=\left(\epsilon_1,...,\epsilon_m\right),
+\end{equation}
+with positive coordinates is called the \textit{reserve vector} and is used as a parameter in the algorithm.
+\item
+For each interval $(x_{i-1},x_i)$,$1 \leq i \leq k+1$, calculate the \textit{characteristic} $R(i)$: 
+\[
+R(i)=\Delta_i+ \frac{(z_i-z_{i-1})^2}{(r_\nu\mu_\nu)^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu=\nu(x_{i-1})=\nu(x_i),
+\]
+\[
+R(i)= 2\Delta_i-4\frac{z_i-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu(x_{i-1})<\nu(x_i)=\nu,
+\]
+\[
+R(i)= 2\Delta_i-4\frac{z_{i-1}-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu = \nu(x_{i-1})>\nu(x_i).
+\]
+where $\Delta_i=(x_i-x_{i-1})^{1/N}$, and the values $r_\nu>1, 1\leq\nu\leq m+1$, are used as parameters in the algorithm.
+\item
+Reorder the characteristics $R(i)$, $1\leq i \leq k+1$, from highest to lowest 	
+\begin{equation}\label{Eq:23}
+R(t_1)\geq R(t_2)\geq ... \geq R(t_{k})\geq R(t_{k+1})
+\end{equation}
+and choose $p$ largest characteristics with interval numbers $t_j, 1\leq j \leq p$.
+\item
+Carry out $p$ new trials in parallel at the points $x^{k+j}, 1 \leq j \leq p$, calculated by the formulae
+\begin{eqnarray*}
+& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}, \; \nu(x_{t_j-1})\neq \nu(x_{t_j}), \\
+& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}- \frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \; \nu(x_{t_j-1})=\nu(x_{t_j})=\nu. \\
+\end{eqnarray*}
+
+%Добавить результаты испытаний в информационную базу алгоритма, и перейти к правилу 1.
+\end{enumerate}
+
+The algorithm stops if the condition $\Delta_{t_j}\leq \epsilon$ becomes true for at least one number $t_j, 1\leq j \leq p$; here  $\epsilon>0$ has an order of magnitude of the desired coordinate accuracy.
+
+A detailed description of the algorithm convergence theory is presented in \cite{Strongin2000}.
+
+\section{Parallel algorithm for mixed-integer problems}
 
@@
 имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_i = i, 1\leq i \leq S-1$.
+%значения функций и индексов в данных точках - не определены!
 Используя введенные обозначения можно переформулировать исходную задачу как
-\[
+\begin{equation}\label{problem_is1}
 \min \left\{g_{m+1}(x): x \in [0,S], \; g_j(x) \leq 0, \; 1 \leq j \leq m\right\}.
-\]
+\end{equation}
 
@@
 
-Applying the global search algorithm to solving the problem (\ref{problem_is1}), we will find 
+Applying the parallel index algorithm to solving the problem (\ref{problem_is1}), we will find 
 the solution of the problem (\ref{problem_i}). In this case the major part of trials will conducted in 
@@
 
-Thus, we have constructed the \textit{Mixed-Integer Global Search Algorithm} (MIGSA) based on the  
+Thus, we have constructed the \textit{Mixed-Integer Parallel Index Algorithm} (MIPIA) based on the  
 reduction of the mixed-integer non-convex optimization problem to the non-convex optimization problem. 
@@
 
-Let us compare the proposed MIGSA method with a genetic algorithm for solving the mixed-integer optimization problems implemented in Matlab Global Optimization Toolbox \cite{Matlab}. In Table 
+Let us compare proposed MIPIA with a genetic algorithm for solving the mixed-integer global optimization problems implemented in Matlab Global Optimization Toolbox \cite{Matlab}. In Table 
 \ref{tab:1}, the numbers of trials required for solving the known test mixed-integer 
@@
 computer with Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM under MS Windows 10. The 
-results of experiments have demonstrated the advantage to the MIGSA method in the number 
+results of experiments have demonstrated the advantage of MIPIA in the number 
 of iterations as well as in the execution time.
@@
 \begin{table}
-	\caption{Comparison of MIGSA and GA}
+	\caption{Comparison of MIPIA and GA}
 	\label{tab:1}

### 2019-03-28 2019/RSD_Integer/RSD2019.tex

@@
 
-\title{Parallel Global Optimization for Non-Convex Mixed-Integer Problems}
+\title{Parallel Global Optimization for Non-Convex Mixed-Integer Problems
+\thanks{This study was supported by the Russian Science Foundation, 
+project No.\,16-11-10150.}}
 \author{Konstantin Barkalov \and Ilya Lebedev %\Letter 
@@
 
-By employing the continuous single-valued Peano curve $y(x)$ mapping the unit interval 
+By employing the continuous single-valued Peano-Hilbert curve $y(x)$ mapping the unit interval 
 $[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}), it is possible to find the 
@@
 \]
-Algorithms for numerical construction of Peano curve approximation (\textit{evolvent}) are 
-considered in \cite{Strongin2013}. Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined 
+Algorithms for numerical construction of Hilbert curve approximation (\textit{evolvent}) are 
+considered in \cite{Strongin2013}. These evolvents are fractals generated by an iterative
+process, that fill in the hypercube $D$ with accuracy $2^{-m}$, where integer $m>0$ is the evolvent construction parameter. 
+%Пример разверток - рисунки: 2D или 3D?
+%Examples of the evolvent with different m in two dimensions are given in Fig. 1.
+%In Fig. 1 approximations of levels 1, 2, 3, and 4 of the Hilbert curve are given in the domain [􀀀1; 1] \in R3.
+
+Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined 
 and computable in the subranges 
@@
 \end{itemize}
-The dyad $ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} $ will be referred to as the \textit{trial 
-outcome}.
+The dyad 
+\begin{equation} \label{trial_result}
+ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} 
+\end{equation}
+will be referred to as the \textit{trial outcome}.
 
-An efficient parallel index algorithm (PIA) for solving the constrained optimization problem 
-(\ref{problem}) has been developed at University of Nizhni Novgorod.
+An efficient parallel index algorithm (PIA) for solving the constrained global optimization problem (\ref{problem}) has been developed at University of Nizhni Novgorod.
 
@@
 \end{equation}
-and match them with the values $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, from (\ref{Eq:16}) , calculated at these points; points $x_0=0$ и $x_{k+1}=1$ are introduced additionally; the values $z_0$ и $z_{k+1}$ are indeterminate.
+and match them with the values $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, from (\ref{trial_result}), calculated at these points; points $x_0=0$ и $x_{k+1}=1$ are introduced additionally; the values $z_0$ и $z_{k+1}$ are indeterminate.
 \item
@@
 \begin{gather}\label{problem_i}
-\min{\left\{ g_{m+1}(y,u):y\in D, \; g_j(y,u)\leq 0, \; 1 \leq j \leq m\right\}},\\
-D=\left\{a_i\leq y_i \leq b_i, \; 1\leq i \leq N \right\}.\nonumber
+\min{\left\{ g_{m+1}(y,u):y\in D, \; g_i(y,u)\leq 0, \; 1 \leq i \leq m\right\}},\\
+D=\left\{a_j\leq y_j \leq b_j, \; 1\leq j \leq N \right\}.\nonumber
 \end{gather}
@@
 \begin{gather}\label{problem_is}
- \min_{s\in\{1,...,S\}}\left(\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_j(y,u_s)\leq 0, \; 1 \leq j \leq m\right\}}\right),\\
-D=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, 1 \leq i\leq N \right\}.\nonumber 
+ \min_{s\in\{1,...,S\}}\left(\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_i(y,u_s)\leq 0, \; 1 \leq i \leq m\right\}}\right),\\
+D=\left\{ a_j\leq y_j \leq b_j, \; 1 \leq j\leq N \right\}.\nonumber 
 \end{gather}
@@
 \[
- \min{\left\{ g_{m+1}(y(x),u_s):x \in [0,1], \; g_j(y(x),u_s)\leq 0, \; 1 \leq j \leq m\right\}}, s\in\{1,...,S\}.
+ \min{\left\{ g_{m+1}(y(x),u_s):x \in [0,1], \; g_i(y(x),u_s)\leq 0, \; 1 \leq i \leq m\right\}}, s\in\{1,...,S\}.
 \]
@@
 \[
-g_j(x) = g_j(Y(x),u_{E(x)+1}), x\in[0,S],
+g_i(x) = g_i(Y(x),u_{E(x)+1}), x\in[0,S],
 \]
-имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_i = i, 1\leq i \leq S-1$.
+имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_k = i, 1\leq i \leq S-1$.
 %значения функций и индексов в данных точках - не определены!
@@
 \begin{equation}\label{problem_is1}
-\min \left\{g_{m+1}(x): x \in [0,S], \; g_j(x) \leq 0, \; 1 \leq j \leq m\right\}.
+\min \left\{g_{m+1}(x): x \in [0,S], \; g_i(x) \leq 0, \; 1 \leq i \leq m\right\}.
 \end{equation}
@@
     \centering
-    \includegraphics[height=2.9cm,width=0.8\textwidth]{fig1.jpg}
+    \includegraphics[width=1.0\textwidth] {fig1.jpg}
     \caption{Reduced mixed-integer global optimization problem}
@@
 
-In order to demonstrate the reliability of the MIGSA method, we have solved two series of 100 
-multiextremal mixed-integer problems each constructed on the basis of modified problems of 
-\textit{Simple} and \textit{Hard} classes generated by GKLS generator. %~\cite{Gaviano}. 
+%In order to demonstrate the reliability of the MIGSA method, we have solved two series of 100 multiextremal mixed-integer problems each constructed on the basis of modified problems of \textit{Simple} and \textit{Hard} classes generated by GKLS generator.
+
+Известным генератором тестовых задач непрерывной многоэкстремальной оптимизации является GKLS-генератор \cite{Gaviano}.
+Он позволяет генерировать functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.).
 This generator of multiextremal functions is often used for the investigations of the global optimization algorithms~ 
 \cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
+%Линии уровня какой-нибудь задачи GKLS? C точками испытаний?
+
+В проведенных экспериментах GKLS-генератор использовался как основа для построения mixed-integer problems. Правила, позволяющие генерировать тестовые задачи подобного вида, заключаются в следующем.
+
+\begin{enumerate}
+	\item С помощью генератора GKLS порождается непрерывная многоэкстремальная функция $\varphi(y), \; y\in D, \; D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\}$.
+Глобальный минимум данной функции достигается в известной точке $y'=(y'_1,...,y'_N)$ и равен $\varphi'=\varphi(y')=-1$.
+	\item Генерируется concave mixed-integer function 
+	\[
+			\mu(y,u) = -2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} \right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right].
+	\]
+	где $y\in D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\} \subset R^N$, a $u\in U = = \left\{ a_j\leq u_j\leq b_j, 1\leq j \leq M \right\} \subset Z^N$; т.е.
+данная функция имеет $N$ непрерывных и $M$ дискретных параметров и достигает минимального значения в точке $(y',b)$.
+	\item Вычисляется коэффициент 
+	\[
+	C = 4 - \min_{y,u} \left\{ \mu(y,u) \right\}, \; y\in D, u \in U.
+	\]
+	Очевидно, что минимум вогнутой функции $\mu(y,u)$ в одной из угловых точек области поиска, поэтому при размерности задачи порядка 10 его вычисление может быть выполнено методом перебора.
+	\item Формируется многоэкстремальная mixed-integer function 
+	\[
+	f(y,u) = \left(\varphi(y) + \sum_{j=1}^M{u_j}\right)\left(C + \mu(y,u)\right).
+	\]
+	По построению $f(y,u)$  будет принимать минимальное значение в точке $(y',b)$.
+
+\end{enumerate}
+
 In the problems generated in our experiments, there were 5 integer  and 3 continuous parameters.
-The accuracy of the search %of the solution in the coordinates 
-was equal to $10^{-2}$. All  
-problems of the series have been solved successfully and for solving the problems based on 
-the \textit{Simple} class 11988 trials in average were required whereas for solving the problems based 
-on the \textit{Hard} class 24750 trials were required.
+The accuracy of the search was equal to $10^{-2}$.
 
-\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No 16-11-10150.
+%All problems of the series have been solved successfully and for solving the problems based on the \textit{Simple} class 11988 trials in average were required whereas for solving the problems based on the \textit{Hard} class 24750 trials were required.
+
+\section{Conclusion}
+
+Заключение
+
+%\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No 16-11-10150.
 
@@
 
-%\bibitem{Gaviano} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
+\bibitem{Gaviano} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
 ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003)

### 2019-03-31 2019/PAVT/slides/slides.tex
PAVT: update slides

@@
 \begin{frame}
-  \begin{center}
-
-  \frametitle{Dimension reduction}
-  Peano-type curve \(y(x)\) allows to reduce the dimension of the original problem:
-  \begin{gather}
-    \lbrace y\in \mathbb{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\} \nonumber \\
-    \min\{\varphi(y): y\in D\}=\min\{\varphi(y(x)): x\in [0,1]\} \nonumber
-  \end{gather}
-  \(y(x)\) is non-smooth function which continuously maps the segment \([0,1]\) to the hypercube \(D\).
-  \begin{figure}[ht]
-    \vspace*{-0.5cm}
-    \subfloat{{\includegraphics[width=.35\textwidth]{peano3d.png} }}
-    \subfloat{{\includegraphics[width=.35\textwidth]{peano2d.png} }}
-  \end{figure}
-\end{center}
+  \frametitle{Methods}
+  \begin{itemize}
+    \item[$\square$] \textbf{Deteriminstic}
+    \begin{itemize}
+      \item Have complicated internal structure in multidimensional case;
+      \item Usually store and use the whole history of trials accumulated during search;
+      \item Require non-trivial parallelization schemes;
+      \item Under some assumptions convergence to the global solution is guaranteed;
+    \end{itemize}
+    \item[$\square$] \textbf{Stochastic}
+    \begin{itemize}
+      \item Have relative simple internal structure;
+      \item Require constant amount of memory to store internal state of some random process or individuals of population;
+      \item In most cases allows wide-scale parallelization by trials or by running many copies of the same method
+      with different seeds in parallel.
+      \item Convergence is guaranteed in probabilistic sense only.
+    \end{itemize}
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Properties of the reduced problem}
-  After applying the Peano-type evolvent \(\varphi(y(x))\) satisfies the uniform H{\"o}lder condition:
-  \begin{displaymath}
-    |\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
-  \end{displaymath}
-  \(\varphi(y(x))\) is non-smooth and has multiple local and \textbf{global} extremums even if \(\varphi(y)\) is unimodal.
-  The latter problem is caused by loss of the information about \(N\)-d neighborhood after the transformation to the \(1\)-d space.
-
-  \begin{figure}[ht]
-    \begin{center}
-
-    \vspace*{-0.5cm}
-    \subfloat{\raisebox{.02\textheight}{\includegraphics[width=.3\textwidth]{parabaloid.png}}}
-    \subfloat{\raisebox{.2\textheight}{\includegraphics[width=.05\textwidth]{arrow.jpg}}}
-    \subfloat{{\includegraphics[width=.45\textwidth]{map_paraboloid.pdf}}}
-  \end{center}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Basic optimization method}
-
-  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
-  \begin{enumerate}
-    \setlength{\itemindent}{.1in}
-    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
-    \item[Step 2.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
-    \item[Step 3.] Choose \(p\) intervals \((x_{t_j-1}, x_{t_j})\) with the greatest characteristics and
-    compute objective \(\varphi(y(x^{k+j}))\) in points chosen using the decision rule \(d\):
-    \begin{displaymath}
-      x^{k+1+j}=d(t)\in (x_{t_j-1}, x_{t_j}),\:j=\overline{1,p}
-    \end{displaymath}
-    \item[Step 4.] If \(x_{t_j}-x_{t_j-1}<\varepsilon\) for one of \(j=\overline{1,p}\), stop the method.
-  \end{enumerate}
-  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
+  \frametitle{Methods}
+  In this work considered the following solvers available in open-source:
+  \begin{itemize}
+    \item[$\square$] \textbf{Deteriminstic}
+    \begin{itemize}
+      \item DIRECT, DIRECT$l$;
+      \item AGS, AGS$l$;
+    \end{itemize}
+    \item[$\square$] \textbf{Stochastic}
+    \begin{itemize}
+      \item Multi Level Single Linkage;
+      \item StoGO;
+      \item Differential Evolution;
+      \item Controlled Random Search;
+      \item Dual Simulated Annealing;
+      \item Ant colony optimization;
+    \end{itemize}
+  \end{itemize}
 \end{frame}
@@
     \begin{column}{0.5\textwidth}
-      Generator GKLS was employed to construct the sets of test problems:
+      Generator GKLS was employed to construct 8 sets of 100 test problems:
       \begin{displaymath}
@@
 \begin{frame}
+  \frametitle{Results of sequential methods}
+
+  \begin{table}
+  \begin{center}
+  \caption{Averaged number of trials executed by sequential methods for solving the test
+  optimization problems}
+  \resizebox{\textwidth}{!}{%
+    \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
+      \hline
+      & AGS & AGS\(l\) & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
+    \hline
+    \(F_{GR}\)     & 193.1 &  \textbf{158.3} & 400.3 & 182.3 & 214.9 & 947.2 & 691.2 & 1257.3 &
+  1336.8 \\
+    \hline
+    GKLS 2d Simple &  254.9 & 217.6 & 510.6 & \textbf{189.0} & 255.2 & 556.8 & 356.3 & 952.2
+  & 1251.5 \\
+    \hline
+    GKLS 2d Hard   &  728.7 & \textbf{488.0} & 844.7 & 985.4 & 1126.7 & 1042.5 & 1637.9 &
+  1041.1 & 2532.2 \\
+    \hline
+    GKLS 3d Simple &  1372.1 & 1195.3 & 4145.8 & \textbf{973.6} & 1477.8 & 4609.2 & 2706.5 &
+  5956.94 & 3856.1 \\
+    \hline
+    GKLS 3d Hard   &  3636.1 & \textbf{1930.5} & 6787.0 & 2298.7 & 3553.3 & 5640.1 & 4708.4 &
+  6914.3 & 7843.2 \\
+    \hline
+    GKLS 4d Simple &  26654.1 & 11095.7 & 37436.8 & \textbf{7824.3} & 15994.1 & 41514.3 &
+  21417.9 & 19157.7 & 59895.4 \\
+    \hline
+    GKLS 4d Hard   &  54536.8 &  \textbf{23167.8} &  73779.3 &  23204.4 &  54489.9 &  80247.2
+  &  68815.5 &  27466.1 &  109328.1  \\
+    \hline
+    GKLS 5d Simple &  29810.0 & 11529.0 & 143575.0 & \textbf{7166.5} & 13970.5 & 52647.6 &
+  34255.3 & 73074.5 & 91580.4 \\
+    \hline
+    GKLS 5d Hard   &  113129.1 & 67652.7 & 165192.8 & \textbf{66327.4} & 164390.6 & 138766.2
+  & 116973.1 & 105496.9 & 155123.8 \\
+    \hline
+  \end{tabular}}
+    \label{tab:trials}
+  \end{center}
+  \end{table}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results of sequential methods}
+  \begin{table}
+  \begin{center}
+  \caption{Number of test optimization problems solved by sequential methods}
+    \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
+      \hline
+      & AGS & AGS\(l\) & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
+    \hline
+    \(F_{GR}\)     &  100 & 100 & 76  & 100 & 100 & 97  & 96  & 96  & 67\\
+    \hline
+    GKLS 2d Simple &  100 & 100 & 85  & 100 & 100 & 100 & 100 & 98  & 90\\
+    \hline
+    GKLS 2d Hard   &  100 & 100 & 74  & 100 & 100 & 100 & 93  & 85  & 77 \\
+    \hline
+    GKLS 3d Simple &  100 & 97  & 75  & 100 & 100 & 100 & 89  & 86  & 44 \\
+    \hline
+    GKLS 3d Hard   &  100  & 99   & 72   & 100  & 99   & 100  & 88   & 77   & 43 \\
+    \hline
+    GKLS 4d Simple &  100 & 100 & 46  & 100 & 100 & 94  & 78  & 59  & 16 \\
+    \hline
+    GKLS 4d Hard   &  100 & 100 & 47  & 99  & 97  & 94  & 72  & 32  & 10  \\
+    \hline
+    GKLS 5d Simple &  100 & 100 & 68  & 100 & 100 & 98  & 100 & 77  & 9  \\
+    \hline
+    GKLS 5d Hard   &  97  & 99  & 42  & 100 & 90  & 79  & 84  & 48  & 8 \\
+    \hline
+    \end{tabular}
+    \label{tab:solved}
+  \end{center}
+  \end{table}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Basic optimization method}
+  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
+  \begin{enumerate}
+    \setlength{\itemindent}{.1in}
+    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
+    \item[Step 2.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
+    \item[Step 3.] Choose \(p\) intervals \((x_{t_j-1}, x_{t_j})\) with the greatest characteristics and
+    compute objective \(\varphi(y(x^{k+j}))\) in points chosen using the decision rule \(d\):
+    \begin{displaymath}
+      x^{k+1+j}=d(t)\in (x_{t_j-1}, x_{t_j}),\:j=\overline{1,p}
+    \end{displaymath}
+    \item[Step 4.] If \(x_{t_j}-x_{t_j-1}<\varepsilon\) for one of \(j=\overline{1,p}\), stop the method.
+  \end{enumerate}
+  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
+\end{frame}
+
+\begin{frame}
+  \begin{center}
+  \frametitle{Dimension reduction}
+  Peano-type curve \(y(x)\) allows to reduce the dimension of the original problem:
+  \begin{gather}
+    \lbrace y\in \mathbb{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\} \nonumber \\
+    \min\{\varphi(y): y\in D\}=\min\{\varphi(y(x)): x\in [0,1]\} \nonumber
+  \end{gather}
+  \(y(x)\) is non-smooth function which continuously maps the segment \([0,1]\) to the hypercube \(D\).
+  \begin{figure}[ht]
+    \vspace*{-0.5cm}
+    \subfloat{{\includegraphics[width=.35\textwidth]{peano3d.png} }}
+    \subfloat{{\includegraphics[width=.35\textwidth]{peano2d.png} }}
+  \end{figure}
+\end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Parallel Globalizer solver}
+  \begin{figure}[ht]
+    \includegraphics[width=0.95\textwidth]{evolvents_parallel.pdf}
+  \end{figure}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Parallel MIDACO solver}
+\end{frame}
+
+\begin{frame}
   \frametitle{Cluster environment}
   \begin{center}
-
   The computational experiments have been carried out on the Lobachevsky supercomputer at  State University of Nizhni Novgorod.
@@
 \end{center}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Evolvents comparison}
-  %\begin{figure}[ht]
-  %  \hspace*{-0.9cm}
-  %  \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_opt_pt_op.pdf} }}
-  %  \subfloat[\(r=5.0\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_same_r_opt_pt_op.pdf} }} \hspace*{4cm}
-  %  \caption{Operating characteristics on GKLS 2d Simple class}
-  %\end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Evolvents comparison}
-  %\begin{figure}[ht]
-  %  \hspace*{-0.9cm}
-  %  \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_opt_pt_op.pdf} }}
-  %  \subfloat[\(r=4.5\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_same_r_opt_pt_op.pdf} }}
-  %  \caption{Operating characteristics on GKLS 3d Simple class}
-  %\end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Choice of evolvent for the parallel algorithm}
 \end{frame}
@@
   \frametitle{Results of applying the parallel algorithm}
-
+  \begin{table}
+    \centering
+    \caption{Averaged numbers of iterations executed by the parallel algorithms for solving the test
+  optimization problems}
+    \label{tab:iterations}
+    \begin{tabular}{cccccccc}
+      \cline{3-8}\noalign{\smallskip}
+      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{Globalizer} & &
+  \multicolumn{2}{c}{MIDACO}   \\
+      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
+      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
+  \textit{Hard}  \\
+      \noalign{\smallskip}\hline
+      I & \textbf{1 cluster node}
+        & \textit{1} &   25270 (100)  & 55180 (99) & & 27645 (98) & 72068 (71)  \\
+      &  & \textit{16} & 1765 (100)  & 3714 (100) & &  1640 (97) & 4304 (70) \\
+      \hline \noalign{\smallskip}
+  II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
+    & \textit{1} & 13056 (100) & 22938 (99) & & 10558 (89) & 27128 (73) \\
+  &   & \textit{16} & 732 (100) & 1759 (100)  &  & 1130 (92) & 2254 (73) \\
+      \hline \noalign{\smallskip}
+  III & \textbf{4 cluster nodes} %\multirow{3}{*}{}
+    & \textit{1}  & 5016 (100) & 12703 (100) & & 5777 (94) & 15980 (75) \\
+  & & \textit{16} & 367 (100) & 776 (100) & & 529 (88) &  1264 (66) \\
+      \hline \noalign{\smallskip}
+  VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
+    & \textit{1}  & 2103 (100) & 5063 (100) & & 2847 (97) & 9853 (83)\\
+  & & \textit{16} & 145 (100)  & 310 (100)  & & 272 (83) & 774 (57)\\
+      \hline \noalign{\smallskip}
+  V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
+    & \textit{1}  & 1155 (100) & 2399 (100) & & 2233 (98) & 6022 (86) \\
+  & & \textit{16} & 76 (100)  & 159 (100)  & & 168 (87) & 393 (53)\\
+      \noalign{\smallskip}\hline
+    \end{tabular}
+  \end{table}
 \end{frame}
@@
   \frametitle{Results of applying the parallel algorithm}
+  \begin{table}
+    \centering
+    \caption{Speedup of parallel computations executed by the parallel algorithms}
+    \label{tab:speedup}
+    \begin{tabular}{cccccccc}
+      \cline{3-8}\noalign{\smallskip}
+      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{Globalizer} & &
+  \multicolumn{2}{c}{MIDACO}   \\
+      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
+      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
+  \textit{Hard}  \\
+      \noalign{\smallskip}\hline
+  I  & \textbf{1 cluster node}  %\multirow{3}{*}{}
+      & \textit{1}   & 25270 (27.3s) & 55180 (61.8s) & & 27645 (32.2s) & 72068 (132.5s)  \\
+    &  & \textit{16} & 14.3(11.7) & 14.9(12.6)  & &  16.9(14.4) & 16.7(14.4) \\
+    \hline \noalign{\smallskip}
+  II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
+    & \textit{1}      &   1.9(1.9) & 2.4(2.2)  & & 2.6(1.7) & 2.7(2.3) \\
+    &   & \textit{16} & 34.5(20.4) & 31.4(18.8) & & 24.5(18.8) & 32.0(29.1) \\
+      \noalign{\smallskip}\hline	\noalign{\smallskip}
+  III  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
+  & \textit{1}      & 5.0(4.6) & 4.3(4.1) & & 4.8(3.9) & 4.5(4.4) \\
+  &   & \textit{16} & 68.8(23.7) & 71.2(25.8) & & 52.3(32.9) & 57.0(50.2) \\
+    \noalign{\smallskip}\hline	\noalign{\smallskip}
+  VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
+  & \textit{1}    & 12.0(8.7) & 10.9(8.3) & & 9.7(8.2)    & 7.3(9.0)    \\
+  & & \textit{16} & 174.1(4.7) & 177.8(5.3) & & 101.6(51.6) & 93.1(84.1) \\
+      \noalign{\smallskip}\hline
+      V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
+      & \textit{1}    & 21.9(11.4)  & 23.0(12.8)  & & 12.4(11.1)  & 12.0(14.8)  \\
+      & & \textit{16} & 333.5(2.7) & 347.9(3.0) & & 164.5(82.0) & 183.2(129.7) \\
+          \noalign{\smallskip}\hline
+    \end{tabular}
+  \end{table}
+\end{frame}
 
+\begin{frame}
+  \frametitle{Results of applying the parallel algorithm}
+  \begin{figure}[ht]
+    \centering
+    \subfloat[$p=1$]{{\includegraphics[width=.5\textwidth]{speedup_midaco_globalizer_p_1.pdf}}\label{fig:speedup_p1}}
+    \subfloat[$p=16$]{{\includegraphics[width=.5\textwidth]{speedup_midaco_globalizer_p_16.pdf}}\label{fig:speedup_p16}}
+    \caption{Speedup demonstrated by the parallel algorithms when solving problems from the GKLS 4d Hard class}
+  \end{figure}
 \end{frame}
+
+\begin{frame}
+  \frametitle{Results of applying the parallel algorithm}
+  \begin{figure}[ht]
+    \centering
+    \subfloat[$p=1$]{{\includegraphics[width=.5\textwidth]{time_midaco_globalizer_p_1.pdf}}\label{fig:time_p1}}
+    \subfloat[$p=16$]{{\includegraphics[width=.5\textwidth]{time_midaco_globalizer_p_16.pdf}}\label{fig:time_p16}}
+    \caption{Averaged execution time of the parallel algorithms when solving problems from the GKLS 4d Hard class}
+  \end{figure}
+\end{frame}
+
+%\begin{frame}
+%  \frametitle{Results of applying the parallel algorithm}
+%\end{frame}
 

### 2019-03-31 2019/RSD_Integer/RSD2019.tex
Добавил текст про GPU реализацию и таблицы с экспериментами.

@@
 
+Наиболее эффективно на GPU можно реализовать операцию вычисления сразу многих значений целевой функции.
+Для этого требуется реализовать на GPU процедуру вычисления значений функции.
+Пересылки данных от CPU к GPU будут минимальные: требуется лишь передать на GPU координаты точек испытаний, и получить обратно значения функций в этих точках.
+Функции, определяющие обработку результатов испытаний в соответствии с алгоритмом и требующие работы с большим объемом накопленной поисковой информацией, могут быть эффективно реализованы на CPU.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=1.0\textwidth] {fig2.jpg}
+    \caption{Схема информационных обменов в GPU-алгоритме}
+    \label{fig:2}
+\end{figure}
+
+Общая схема организации вычислений с использованием GPU приведена на Fig.
+\ref{fig:2}.
+В соответствии с данной схемой на CPU выполняются шаг 1 – 6 параллельного алгоритма глобального поиска.
+Координаты \textit{p} точек испытаний, вычисленные на шаге 7 алгоритма, накапливаются в промежуточном буфере, а затем передаются на графический процессор.
+На GPU происходит вычисление значений функции в этих точках, после чего результаты испытаний (снова через промежуточный буфер) передаются на CPU.
+
 \section{Results of experiments}
@@
 
-In the problems generated in our experiments, there were 5 integer  and 3 continuous parameters.
+In the problems generated in our experiments, there were 5 integer  and 6 continuous parameters.
+Дискретные параметры имеют по 4 значения, таким образом имеем \textit{S} = 1024 возможных значения.
 The accuracy of the search was equal to $10^{-2}$.
+С помощью GKLS-генератора создавалось 100 шестимерных задач.
+Computational experiments were carried out on Lobachevsky supercomputer. The node of 
+supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The 
+CPU had 8 cores, i.e. each node had a total of 16 cores and two NVIDIA Kepler K20Х GPUs.
+
+В Table \ref{tab:2} приведены результаты экспериментов на CPU с использованием OpenMP.
+Использовалось 1, 8 и 16 потоков.
+16 потоков соответствуют полной загрузке CPU на узле кластера.
+Приведено среднее число итераций требуемое для решения задачи, среднее время решения и ускорение по времени по отношению к последовательному запуску.
+За одну итерации проводилось столько же испытаний сколько использовалось потоков.
+
+\begin{table}
+	\caption{Результаты экспериментов на CPU}
+	\label{tab:2}
+	\center
+	\begin{tabular}{cccccc}
+		\hline\noalign{\smallskip}
+
+		Число потоков & \textit{p} = 1 & \textit{p} = 8 & \textit{p} = 16 \\
+	\noalign{\smallskip} \hline \noalign{\smallskip}
+
+		Среднее число итераций  & 3 221 023,4 & 512 314,2 & 209 237,0 \\
+		Среднее время решения & 5 520,3 & 1 253,5 & 717,3 \\
+		Ускорение по отношению к одному потоку & 1,0 & 4,4 & 7,7 \\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+Наблюдается почти линейное ускорение по итерациям и значительное ускорение по времени решения задачи.
+
+В Table \ref{tab:3} приведены результаты экспериментов на GPU с использованием CUDA.
+Использовалось 256, 512 и 1024 нитей.
+Приведено среднее число итераций требуемое для решения задачи, среднее время решения и ускорение по времени по отношению к полной загрузке CPU на узле кластера.
+За одну итерации проводилось столько же испытаний сколько использовалось нитей.
+
+\begin{table}
+	\caption{Результаты экспериментов на GPU}
+	\label{tab:3}
+	\center
+	\begin{tabular}{cccccc}
+		\hline\noalign{\smallskip}
+
+		Число нитей & \textit{p} = 256 & \textit{p} = 512 & \textit{p} = 1024 \\
+	\noalign{\smallskip} \hline \noalign{\smallskip}
+
+		Среднее число итераций  & 1 522,3 & 919,8 & 412,6 \\
+		среднее время решения & 33,6 & 31,2 & 30,2 \\
+		Ускорение по отношению к полной загрузке ЦПУ & 21,4 & 23,0 & 23,7 \\
+
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+Наблюдается значительное ускорение. Ускорение по отношению к последовательному запуску составило 182.
 

### 2019-03-31 2019/PAVT/slides/slides.tex
PAVT: the final version of the slides

@@
   \begin{itemize}
-    \item[$\square$] \textbf{Deteriminstic}
+    \item[$\square$] \textbf{Deteriminstic}:
     \begin{itemize}
@@
       \item Require non-trivial parallelization schemes;
-      \item Under some assumptions convergence to the global solution is guaranteed;
+      \item Under some assumptions convergence to the global solution is guaranteed.
     \end{itemize}
-    \item[$\square$] \textbf{Stochastic}
+    \item[$\square$] \textbf{Stochastic}:
     \begin{itemize}
@@
   \begin{itemize}
-    \item[$\square$] \textbf{Deteriminstic}
+    \item[$\square$] \textbf{Deteriminstic}:
     \begin{itemize}
       \item DIRECT, DIRECT$l$;
-      \item AGS, AGS$l$;
+      \item AGS, AGS$l$.
     \end{itemize}
-    \item[$\square$] \textbf{Stochastic}
+    \item[$\square$] \textbf{Stochastic}:
     \begin{itemize}
@@
       \item Dual Simulated Annealing;
-      \item Ant colony optimization;
+      \item Ant Colony Optimization.
     \end{itemize}
@@
 \begin{frame}
-  \frametitle{Basic optimization method}
+  \frametitle{Univariate Algorithm of Global Search}
   Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
@@
 \begin{frame}
-  \frametitle{Parallel Globalizer solver}
+  \frametitle{Globalizer parallel solver}
   \begin{figure}[ht]
@@
 \begin{frame}
-  \frametitle{Parallel MIDACO solver}
+  \frametitle{MIDACO parallel solver}
+  The MIDACO solver utilizes reverse communication architecture.
+  Reverse communication means that the call of the objective function happens outside and
+  independently of the MIDACO source code.
+  Relying on this feature the following distributed parallelization scheme was implemented:
+  \begin{enumerate}
+    \item MIDACO runs at the master node and generates \(n \times p\) new trial points at each
+  iteration.
+    \item Each of \(n\) distributed computational nodes gets \(p\) points from the master and performs
+  \(p\) parallel trials on local computing devices.
+    \item Each of \(n\) distributed computational nodes sends \(p\) values of the objective function to
+  the master.
+  \end{enumerate}
+
+  In this scheme all distributed communications were implemented using the MPI library.
+  Parallelism within a single node is powered by the OpenMP standard.
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Results of applying the parallel algorithm}
+  \frametitle{Results of the parallel algorithms}
   \begin{table}
@@
 \begin{frame}
-  \frametitle{Results of applying the parallel algorithm}
+  \frametitle{Results of the parallel algorithms}
   \begin{figure}[ht]
@@
 
-%\begin{frame}
-%  \frametitle{Results of applying the parallel algorithm}
-%\end{frame}
-
 \begin{frame}
@@
     \begin{itemize}
-      \item 1
+      \item AGS\(l\) method implemented in the Globalizer system has demonstrated the convergence
+    speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-source
+    implementations of which are available;
+    \item the stochastic optimization methods inferior to the deterministic ones in the convergence
+  speed and in reliability. It is manifested especially strongly on more complex multiextremal
+  problems;
+    \item the parallel version of the Globalizer system demonstrates good speedup
+  when running on several nodes, on each of which a single objective function value per iterations is
+  computed. When solving the problems with fast computable objective functions and using multiple
+  threads on the nodes, the speedup for the Globalizer system degrade with increasing
+  number of nodes;
+    \item the MIDACO system is the most suitable for simple global optimization problems with the
+  fast-computable objective functions. In this case, MIDACO is reliable enough and provides a linear
+  speedup with increasing number of nodes and threads executed on these ones in parallel.
     \end{itemize}

### 2019-04-03 2019/RSD_Integer/RSD2019.tex
Поправлен текст. Заменена картинка, добавлен исходный документ с картинкой.

@@
 Дискретные параметры имеют по 4 значения, таким образом имеем \textit{S} = 1024 возможных значения.
+С помощью GKLS-генератора создано 100 шестимерных задач, которые были преобразованы в смешанные 11-ти мерные задачи.
 The accuracy of the search was equal to $10^{-2}$.
-С помощью GKLS-генератора создавалось 100 шестимерных задач.
 Computational experiments were carried out on Lobachevsky supercomputer. The node of 
@@
 Использовалось 1, 8 и 16 потоков.
-16 потоков соответствуют полной загрузке CPU на узле кластера.
-Приведено среднее число итераций требуемое для решения задачи, среднее время решения и ускорение по времени по отношению к последовательному запуску.
-За одну итерации проводилось столько же испытаний сколько использовалось потоков.
+Приведено среднее число итераций требуемое для решения задачи, среднее время решения (в секундах) и ускорение по времени по отношению к последовательному запуску.
+За одну итерацию проводилось столько же испытаний сколько использовалось потоков.
 
@@
 
-Наблюдается почти линейное ускорение по итерациям и значительное ускорение по времени решения задачи.
+Как видно из Table \ref{tab:2} , наблюдается почти линейное ускорение по итерациям и значительное ускорение по времени решения задачи.
+В среднем при последовательном решение задачи требовалось 1,5 часа.
 
@@
 Использовалось 256, 512 и 1024 нитей.
-Приведено среднее число итераций требуемое для решения задачи, среднее время решения и ускорение по времени по отношению к полной загрузке CPU на узле кластера.
+Приведено среднее число итераций требуемое для решения задачи, среднее время решения (в секундах) и ускорение по времени по отношению к полной загрузке CPU на узле кластера.
 За одну итерации проводилось столько же испытаний сколько использовалось нитей.
@@
 
-Наблюдается значительное ускорение. Ускорение по отношению к последовательному запуску составило 182.
+Наблюдается значительное ускорение как по числу итераций, так и по времени.
+В среднем решение задачи (при 1024 нитях) потребовало 30 секунд.
+Таким образом, ускорение по отношению к последовательному запуску составило 182.
 

### 2019-04-03 2019/PAVT/slides/slides.tex
PAVT: camera-ready slides

@@
       \item In most cases allows wide-scale parallelization by trials or by running many copies of the same method
-      with different seeds in parallel.
+      with different seeds in parallel;
       \item Convergence is guaranteed in probabilistic sense only.
@@
       \begin{itemize}
-        \item the number of local minimas;
+        \item the number of local minimums;
         \item the size of the global minima attraction region;
@@
   \end{columns}
+  Also one set of 100 pre-defined highly multiextremal functions (with 10-30 local minimums) was used.
 \end{frame}
@@
   \end{table}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Parallel optimization methods}
+  \begin{itemize}
+    \item Parallel AGS (Globalizer system);
+    \item Parallel ACO (MIDACO system).
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Results of the parallel algorithms}
+  \frametitle{Results of parallel algorithms}
   \begin{table}
@@
 \begin{frame}
-  \frametitle{Results of the parallel algorithms}
+  \frametitle{Results of parallel algorithms}
   \begin{figure}[ht]

### 2019-04-04 2019/RSD_Integer/RSD2019.tex

--- 2019-03-28 2019/RSD_Integer/RSD2019.tex
@@
 \\
-\email{konstantin.barkalov@itmm.unn.ru}}
+\email{ \{konstantin.barkalov,ilya.lebedev\}@itmm.unn.ru}}
 
@@
 \begin{abstract}
+
 In the present paper, the mixed-integer global optimization problems are considered. A novel 
-deterministic algorithm for solving the problems of this class based on the information-statistical 
-approach to solving the continuous global optimization problems has been proposed. The 
-comparison of this algorithm with known analogs demonstrating the efficiency of the 
-developed approach has been conducted. The stable operation of the algorithm was confirmed 
-also by solving a series of several hundred mixed-integer global optimization problems.
+parallel algorithm for solving the problems of this class based on the information-statistical 
+approach to solving the continuous global optimization problems has been proposed. 
+The comparison of this algorithm with known analogs demonstrating the efficiency of the 
+developed approach has been conducted. 
+%Russian
+Предложенный алгоритм допускает эффективное использование graphics accelerators. 
+Проведены вычислительные эксперименты при решении серии из 100 multiextremal mixed-integer problems, которые подтверждают хорошее ускорение алгоритма с использованием GPU.
 
@@
 of the whole search domain. 
-%As a result, the search of the global optimum is reduced to the construction of a coverage (in general, a nonuniform one) of the space of parameters. 
+As a result, the search of the global optimum is reduced to the construction of a coverage (in general, a nonuniform one) of the space of parameters. 
 The problems, in which some parameters can take the integer values only (mixed-integer global 
@@
 
+The situation when some parameters are featured by the discreteness or integerness is frequent in applied problems.
+As a rule, the integer parameters take a small number of values and may denote, for example, the trademarks of the materials used, the variant of typical layouts of components, etc.
+
 A lot of publication have been devoted to the methods of solving the mixed-integer problems 
@@
 methods of solving the problems of this class are based, as a rule, on the Branch-and-Bound 
-approach or on the Branch-and-Reduce one. Also, a number of the metaheuristic and genetic 
-algorithms are known, which are based one way or another on the random search concept.
-%  based ones, anyway, based on the ideas of random search.
+approach \cite{Belotti} or on the Branch-and-Reduce one \cite{Vigerske}. Also, a number of the metaheuristic and genetic 
+algorithms are known, which are based one way or another on the random search concept \cite{Deep,Schluter}.
 
-In the present study, we proposed a novel deterministic method for solving the mixed-integer 
+In the present study, we proposed a novel parallel method for solving the mixed-integer 
 problems based on the information-statistical approach to solving the global optimization 
-problems \cite{Strongin2000,Strongin2013}. The paper text reflecting the results of preliminary 
-studies is composed as follows. First, an approach to solving the problems with continuous 
-parameters has been considered, then its generalization for the mixed-integer problems was 
-proposed. In the final section, the results of comparison of the proposed method with the 
-known analogs are presented.
+problems \cite{Strongin2000,Strongin2013}. 
+%Russian
+В рамках данного подхода 
+\begin{itemize}
+	\item 
+	решение многомерных задач сводится  к решению эквивалентных им одномерных, cоответствующая редукция основана на использовании space-filling curves;
+	\item 
+	при решении constrained optimization problems каждое ограничение учитывается и обрабатывается отдельно, штрафные функции не используются;
+	\item 
+	распараллеливание процесса поиска осуществляется за счет одновременного вычисления на каждой итерации нескольких значений целевой функции в разных точках search domain.	
+\end{itemize}
+
+Текст статьи, отражающей результаты проведенного исследования, построен следующим образом. 
+В разделе 2 дано краткое описание описание схемы редукции размерности с использованием space-filling curves, а также описана индексная схема учета ограничений.
+Здесь же приведена формулировка параллельного индексного алгоритма для решения задач непрерывной глобальной оптимизации.
+В разделе 3 предложен подход к обобщению параллельного индексного алгоритма с целью решения mixed-integer problems.
+Приведен способ, позволяющий свести решение mixed-integer problem к решению набора задач непрерывной оптимизации, которое может выполняться параллельно.
+Раздел 4 содержит результаты численных экспериментов.
+Здесь производится сравнение последовательной версии алгоритма с известными аналогами, а также показана эффективность параллельных CPU- и GPU-версий алгоритма при решении серии multiextremal mixed-integer problem.
+Section 5 concludes the paper.
 
@@
 process, that fill in the hypercube $D$ with accuracy $2^{-m}$, where integer $m>0$ is the evolvent construction parameter.
-%Пример разверток - рисунки: 2D или 3D?
-%Examples of the evolvent with different m in two dimensions are given in Fig. 1.
-%In Fig. 1 approximations of levels 1, 2, 3, and 4 of the Hilbert curve are given in the domain [􀀀1; 1] \in R3.
+
+Problems of numerical construction of Peano-type space filling curves and the corresponding theory are considered in detail in \cite{Strongin2000,Strongin2013}.
+Examples of the evolvent with different $m$ in two dimensions are given in Fig.~\ref{evolvents}.
+
+\begin{figure}
+\begin{minipage}{0.49\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (b)}
+\end{minipage}
+\caption{Evolvents in two dimensions with (a) $m=4$ and (b) $m=5$}
+\label{evolvents}
+\end{figure}
 
@@
 An efficient parallel index algorithm (PIA) for solving the constrained global optimization problem (\ref{problem}) has been developed at University of Nizhni Novgorod.
+The scheme of the algorithm is as follows.
 
@@
 
-%Добавить результаты испытаний в информационную базу алгоритма, и перейти к правилу 1.
 \end{enumerate}
@@
 The algorithm stops if the condition $\Delta_{t_j}\leq \epsilon$ becomes true for at least one number $t_j, 1\leq j \leq p$; here  $\epsilon>0$ has an order of magnitude of the desired coordinate accuracy.
-
 A detailed description of the algorithm convergence theory is presented in \cite{Strongin2000}.
@@
 \section{Parallel algorithm for mixed-integer problems}
-
+%Russian
 Рассмотрим теперь случай, когда аргумент функций задачи содержит две компоненты: вектор $y$, принадлежащий гиперинтервалу $D$, и вектор $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е.
@@
 имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_k = i, 1\leq i \leq S-1$.
-%значения функций и индексов в данных точках - не определены!
+Значение  $z_k = g_\nu(y(x_k))$ из (\ref{trial_result}) в этих точках будем считать неопределенным, а значения индекса -- равным 0, т.е. $\nu(x_k) = 0$ .
+
 Используя введенные обозначения можно переформулировать исходную задачу как
@@
 reduction of the mixed-integer non-convex optimization problem to the non-convex optimization problem. 
-%It is possible to prove that the convergence conditions for this algorithm will follow from the convergence conditions of its prototype (GSA).
+%Russian
+Предложенная схема основана на параллельном индексном методе для решения задач непрерывной оптимизации и не является ориентированной на конкретное вычислительное устройство.
+Одной из ключевых операций здесь является параллельное проведение нескольких испытаний в различных точках области поиска (см.
+шаг 7 алгоритма), которе может быть реализовано как на CPU (с использованием OpenMP и/или MPI), так и на GPU (с использованием CUDA).
+Вопросы использования GPU для этих целей были детально рассмотрены в \cite{Barkalov2016,Gergel2016}, указанный подход к использованию GPU был применен и при реализации алгоритма MIPIA.
 
 \section{Results of experiments}
+%Russian
+Первая серия экспериментов была проведена с использованием последовательной версии предложенного алгоритма с целью его сравнения с известными методами аналогичного назначения.
 
@@
 \ref{tab:1}, the numbers of trials required for solving the known test mixed-integer 
-problems by these methods are presented. For both methods, the same accuracy of search $10^{-2}$ were used. All numerical experiments were conducted on a 
+problems by these methods are presented. For both methods, the same accuracy of search $10^{-2}$ were used. These numerical experiments were conducted on a 
 computer with Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM under MS Windows 10. The 
@@
 		 Problem 2 \cite{Floudas}&	481 &	0.0601 & &	417 &	0.04 \\
-%		 Problem 3 \cite{Floudas}& 	1821 &	0.1130 & & 3324 &	0.107 \\
+		 Problem 3 \cite{Floudas}& 	1821 &	0.1130 & & 3324 &	0.107 \\
 		 Problem 6 \cite{Floudas}&	641 &	0.0510 & &	118 &	0.001 \\
@@
 
-%In order to demonstrate the reliability of the MIGSA method, we have solved two series of 100 multiextremal mixed-integer problems each constructed on the basis of modified problems of \textit{Simple} and \textit{Hard} classes generated by GKLS generator.
+Последующие серии экспериментов были выполнены для оценки ускорения параллельной версии предложенного алгоритма MIPIA с использованием как CPU, так и GPU.
+В данных экспериментах решалась серия из 100 тестовых mixed-integer problems, сгенерированных случайных образом.
 
@@
 \cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}. 
-%Линии уровня какой-нибудь задачи GKLS? C точками испытаний?
+In Fig.~\ref{example} (a) and (b), the contour plots of two-dimensional GKLS functions are presented.
+Figures also shows the points of the trials performed by the method until the required accuracy $\epsilon=10^{-2}$ was achieved.
+
+\begin{figure}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{GKLS-4.png} \\ (a)}
+\end{minipage}
+\hfill
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{GKLS-6.png} \\ (b)}
+\end{minipage}
+\caption{Solving a two-dimensional problem using the index algorithm}
+\label{example}
+\end{figure}
 
@@
 	\[
-			\mu(y,u) = -2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} \right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right].
+			\mu(y,u) = -2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} \right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right],
 	\]
-	где $y\in D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\} \subset R^N$, a $u\in U = = \left\{ a_j\leq u_j\leq b_j, 1\leq j \leq M \right\} \subset Z^N$; т.е.
-данная функция имеет $N$ непрерывных и $M$ дискретных параметров и достигает минимального значения в точке $(y',b)$.
+	где 
+	\begin{gather}
+	y\in D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\} \subset R^N,\nonumber \\
+	u\in U = \left\{ u_j \in  \left\{a_j, ..., b_j \right\}, 1\leq j \leq M \right\};\nonumber
+	\end{gather}
+	т.е. данная функция имеет $N$ непрерывных и $M$ дискретных параметров и достигает минимального значения в точке $(y',b)$.
 	\item Вычисляется коэффициент 
@@
 
-In the problems generated in our experiments, there were 5 integer  and 3 continuous parameters.
-The accuracy of the search was equal to $10^{-2}$.
+In the problems generated in our experiments, there were 5 discrete and 6 continuous parameters 
+	\begin{gather}
+	y\in D = \left\{ -1 \leq y_j\leq 1, 1\leq j \leq 6 \right\} \subset R^6,\nonumber \\
+	u\in U = \left\{ u_j \in  \left\{-1, -1/3, 0, 1/3, 1 \right\}, 1\leq j \leq 5 \right\}.\nonumber
+	\end{gather}
 
-%All problems of the series have been solved successfully and for solving the problems based on the \textit{Simple} class 11988 trials in average were required whereas for solving the problems based on the \textit{Hard} class 24750 trials were required.
+Было сгенерировано 100 11-и мерных mixed-integer problems данного типа. The accuracy of the search was equal to $10^{-2}$.   
+Computational experiments were carried out on Lobachevsky supercomputer. The node of 
+supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The 
+CPU had 8 cores, i.e. each node had a total of 16 cores and two NVIDIA Kepler K20Х GPUs.
+
+В Table \ref{tab:2} приведены результаты экспериментов на CPU с использованием OpenMP в зависимости от числа задействованных threads $p$.
+Использовалось 1, 8 и 16 потоков.
+Приведено среднее число итераций $K_{av}$, требуемое для решения задачи, среднее время решения $T_{av}$ (в секундах), time speedup $S$ and iteration speedup $s$ (по отношению к последовательному запуску, т.е.
+$p=1$).
+В соответствии со схемой распараллеливания число испытаний на одной итерации параллельного алгоритма равнялось числу задействованных threads.
+
+\begin{table}
+	\caption{Результаты экспериментов на CPU}
+	\label{tab:2}
+	\center
+	\begin{tabular}{cccccc}
+		\hline\noalign{\smallskip}        
+		$p$ & $T_{av}$ & $K_{av}$ & $S$ & $s$ \\
+	\noalign{\smallskip} \hline \noalign{\smallskip}
+	1 \;&	5520 \;  & 3221023 \;  & ...\; & ...  \\
+	8 \;&	1253 \; & 512314 \;  & 4.4\; & 6.3  \\
+	16\;&  717 \; &  209237 \; & 7.7\; & 15.4 \\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+Как видно из Table \ref{tab:2} , наблюдается почти линейное ускорение по итерациям и значительное ускорение по времени решения задачи, при этом последовательному алгоритму в среднем требовалось примерно 1.5 часа для решения задачи.
+
+В Table \ref{tab:3} приведены результаты экспериментов на GPU с использованием CUDA в зависимости от числа задействованных GPU-threads.
+Приведено среднее число итераций $K_{av}$, требуемое для решения задачи, среднее время решения $T_{av}$ (в секундах), time speedup $S$ and iteration speedup $s$ (по отношению к полной загрузке CPU на узле кластера).
+
+\begin{table}
+	\caption{Результаты экспериментов на GPU}
+	\label{tab:3}
+	\center
+	\begin{tabular}{cccccc}
+		\hline\noalign{\smallskip}
+			$p$ & $T_{av}$ & $K_{av}$ & $S$ & $s$ \\
+	\noalign{\smallskip} \hline \noalign{\smallskip}
+	256 \;&	33.6 \; & 1522\;  & 21.3 \; & 137  \\
+	512 \;&	31.2 \; & 919 \;  & 23.0 \; & 228  \\
+	1024\;& 30.2 \; & 412 \; & 23.7  \; & 508  \\
+	2048\;& 38.5 \; & 244 \; & 18.6  \; & 858  \\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
+
+Результаты эксперимнетов показывают почти одинаковое time speedup и линейное iteration speedup при использовании $p=256, 512, 1024$ threads.
+Однако при использовании $p=2048$ threads время работы алгоритма замедляется, но число итераций продолжает падать.
+Данный эффект объясняется тем, что для параллельного проведения $p$ испытаний используется GPU, а для обработки результатов испытаний (которая подразумевает работу со всей накопленной на предыдущих итерациях поисковой информацией) используется CPU.
+И при использовании $p=2048$ GPU threads время, которое тратится на передачу и обработку результатов 2048 испытаний, становится сопоставимым со временем проведения испытания, что и приводит к замедлению алгоритма в целом.
 
 \section{Conclusion}
-
-Заключение
-
-%\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No 16-11-10150.
+%Russian
+В данной статье представлены результаты, полученные в Lobachevsky State University of Nizhni Novgorod при разработке и  исследовании параллельных алгоритмов глобальной оптимизации для решения многоэкстремальных задач, в которых часть параметров является непрерывными, а часть - дискретными.
+Для задач указанного класса был предложен эффективный параллельный алгоритм их решения алгоритм. 
+В последовательном варианте данный алгоритм не уступает алгоритму аналогичного назначения, реализованному в Matlab Global Optimization Toolbox.
+В параллельном варианте предложенный алгоритм допускает эффективную реализацию как на CPU, так и на GPU. 
+Проведены вычислительные эксперименты на серии из 100 multiextremal mixed-integer problems, убедительно демонстрирующие хорошее ускорение алгоритма с использованием GPU.
+Так, параллельный алгоритм с использованием GPU показал time speedup $S = 182$ по отношению к последовательному алгоритму и $S = 23.7$ по отношению к алгоритму, полностью задействующему два CPU на узле кластера.
 
@@
 
+\bibitem{Deep}
+Deep, K., Singh, K. P., Kansal, M.L., Mohan, C.: A real coded genetic algorithm for solving integer and mixed integer optimization problems. Appl. Math. Comput. \textbf{212}(2), 505--518 (2009)
+
+\bibitem{Schluter}
+Schl\"uter, M., Egea, J.A., Banga, J.R.: Extended ant colony optimization for non-convex mixed integer nonlinear programming. Computers and Operations Research \textbf{36}(7), 2217--2229 (2009)
+
+\bibitem{Belotti}
+Belotti, P., Lee, J., Liberti, L., Margot, F., W\"achter, A.: Branching and bounds tightening techniques for non-convex MINLP. Optim. Method. Softw. \textbf{24}(4-5), 597--634 (2009)
+
+\bibitem{Vigerske}
+Vigerske, S., Gleixner, A.: SCIP: global optimization of mixed-integer nonlinear programs in a branch-and-cut framework. 
+Optim. Method. Softw. \textbf{33}(3), 563--593 (2018)
+
 \bibitem{Strongin2000}
@@
 https://www.mathworks.com/help/gads/mixed-integer-optimization.html
-
-\bibitem{Deep}
-Deep, K., Singh, K. P., Kansal, M.L., Mohan, C.: A real coded genetic algorithm for solving integer and mixed integer optimization problems. Appl. Math. Comput. \textbf{212}(2), 505--518 (2009)
 
@@
 
+\bibitem{Barkalov2016} 
+Barkalov, K., Gergel, V.: Parallel global optimization on GPU. Journal of Global Optimization 66(1), 2--20 (2016)
+
+\bibitem{Gergel2016}
+Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster. In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings, vol. 1738, art. no. 400006 (2016)
+
 \end{thebibliography}

### 2019-04-07 2019/MOTOR_Sovrasov/slides/slides.tex
MOTOR: add initial version of slides

--- 2019-04-03 2019/PAVT/slides/slides.tex
@@
 
-\title{Comparison Of Several Sequential And Parallel Derivative-free Global Optimization Algorithms}
-\author{\underline{\textbf{Vladislav~Sovrasov}} and \textbf{Semen~Bevzuk}}
+\title{Comparison of Several Stochastic and Deterministic Derivative-free Global Optimization Algorithms}
+\author{\textbf{Vladislav~Sovrasov}}
 \institute{Lobachevsky State University of Nizhni Novgorod}
@@
 \begin{frame}
-  \frametitle{Globalizer parallel solver}
-  \begin{figure}[ht]
-    \includegraphics[width=0.95\textwidth]{evolvents_parallel.pdf}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{MIDACO parallel solver}
-  The MIDACO solver utilizes reverse communication architecture.
-  Reverse communication means that the call of the objective function happens outside and
-  independently of the MIDACO source code.
-  Relying on this feature the following distributed parallelization scheme was implemented:
-  \begin{enumerate}
-    \item MIDACO runs at the master node and generates \(n \times p\) new trial points at each
-  iteration.
-    \item Each of \(n\) distributed computational nodes gets \(p\) points from the master and performs
-  \(p\) parallel trials on local computing devices.
-    \item Each of \(n\) distributed computational nodes sends \(p\) values of the objective function to
-  the master.
-  \end{enumerate}
-
-  In this scheme all distributed communications were implemented using the MPI library.
-  Parallelism within a single node is powered by the OpenMP standard.
-\end{frame}
-
-\begin{frame}
-  \frametitle{Cluster environment}
-  \begin{center}
-  The computational experiments have been carried out on the Lobachevsky supercomputer at  State University of Nizhni Novgorod.
-  One node includes 2 Intel Sandy Bridge E5-2660 2.2 GHz processors and 64 GB RAM.
-
-  Each node runs under CentOS 7 Linux with GCC 4.8 compiler and Intel MPI library.
-\end{center}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results of the parallel algorithms}
-  \begin{table}
-    \centering
-    \caption{Averaged numbers of iterations executed by the parallel algorithms for solving the test
-  optimization problems}
-    \label{tab:iterations}
-    \begin{tabular}{cccccccc}
-      \cline{3-8}\noalign{\smallskip}
-      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{Globalizer} & &
-  \multicolumn{2}{c}{MIDACO}   \\
-      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
-      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
-  \textit{Hard}  \\
-      \noalign{\smallskip}\hline
-      I & \textbf{1 cluster node}
-        & \textit{1} &   25270 (100)  & 55180 (99) & & 27645 (98) & 72068 (71)  \\
-      &  & \textit{16} & 1765 (100)  & 3714 (100) & &  1640 (97) & 4304 (70) \\
-      \hline \noalign{\smallskip}
-  II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
-    & \textit{1} & 13056 (100) & 22938 (99) & & 10558 (89) & 27128 (73) \\
-  &   & \textit{16} & 732 (100) & 1759 (100)  &  & 1130 (92) & 2254 (73) \\
-      \hline \noalign{\smallskip}
-  III & \textbf{4 cluster nodes} %\multirow{3}{*}{}
-    & \textit{1}  & 5016 (100) & 12703 (100) & & 5777 (94) & 15980 (75) \\
-  & & \textit{16} & 367 (100) & 776 (100) & & 529 (88) &  1264 (66) \\
-      \hline \noalign{\smallskip}
-  VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-    & \textit{1}  & 2103 (100) & 5063 (100) & & 2847 (97) & 9853 (83)\\
-  & & \textit{16} & 145 (100)  & 310 (100)  & & 272 (83) & 774 (57)\\
-      \hline \noalign{\smallskip}
-  V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
-    & \textit{1}  & 1155 (100) & 2399 (100) & & 2233 (98) & 6022 (86) \\
-  & & \textit{16} & 76 (100)  & 159 (100)  & & 168 (87) & 393 (53)\\
-      \noalign{\smallskip}\hline
-    \end{tabular}
-  \end{table}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results of parallel algorithms}
-  \begin{table}
-    \centering
-    \caption{Speedup of parallel computations executed by the parallel algorithms}
-    \label{tab:speedup}
-    \begin{tabular}{cccccccc}
-      \cline{3-8}\noalign{\smallskip}
-      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{Globalizer} & &
-  \multicolumn{2}{c}{MIDACO}   \\
-      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
-      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
-  \textit{Hard}  \\
-      \noalign{\smallskip}\hline
-  I  & \textbf{1 cluster node}  %\multirow{3}{*}{}
-      & \textit{1}   & 25270 (27.3s) & 55180 (61.8s) & & 27645 (32.2s) & 72068 (132.5s)  \\
-    &  & \textit{16} & 14.3(11.7) & 14.9(12.6)  & &  16.9(14.4) & 16.7(14.4) \\
-    \hline \noalign{\smallskip}
-  II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
-    & \textit{1}      &   1.9(1.9) & 2.4(2.2)  & & 2.6(1.7) & 2.7(2.3) \\
-    &   & \textit{16} & 34.5(20.4) & 31.4(18.8) & & 24.5(18.8) & 32.0(29.1) \\
-      \noalign{\smallskip}\hline	\noalign{\smallskip}
-  III  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
-  & \textit{1}      & 5.0(4.6) & 4.3(4.1) & & 4.8(3.9) & 4.5(4.4) \\
-  &   & \textit{16} & 68.8(23.7) & 71.2(25.8) & & 52.3(32.9) & 57.0(50.2) \\
-    \noalign{\smallskip}\hline	\noalign{\smallskip}
-  VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-  & \textit{1}    & 12.0(8.7) & 10.9(8.3) & & 9.7(8.2)    & 7.3(9.0)    \\
-  & & \textit{16} & 174.1(4.7) & 177.8(5.3) & & 101.6(51.6) & 93.1(84.1) \\
-      \noalign{\smallskip}\hline
-      V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
-      & \textit{1}    & 21.9(11.4)  & 23.0(12.8)  & & 12.4(11.1)  & 12.0(14.8)  \\
-      & & \textit{16} & 333.5(2.7) & 347.9(3.0) & & 164.5(82.0) & 183.2(129.7) \\
-          \noalign{\smallskip}\hline
-    \end{tabular}
-  \end{table}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results of parallel algorithms}
-  \begin{figure}[ht]
-    \centering
-    \subfloat[$p=1$]{{\includegraphics[width=.5\textwidth]{speedup_midaco_globalizer_p_1.pdf}}\label{fig:speedup_p1}}
-    \subfloat[$p=16$]{{\includegraphics[width=.5\textwidth]{speedup_midaco_globalizer_p_16.pdf}}\label{fig:speedup_p16}}
-    \caption{Speedup demonstrated by the parallel algorithms when solving problems from the GKLS 4d Hard class}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results of parallel algorithms}
-  \begin{figure}[ht]
-    \centering
-    \subfloat[$p=1$]{{\includegraphics[width=.5\textwidth]{time_midaco_globalizer_p_1.pdf}}\label{fig:time_p1}}
-    \subfloat[$p=16$]{{\includegraphics[width=.5\textwidth]{time_midaco_globalizer_p_16.pdf}}\label{fig:time_p16}}
-    \caption{Averaged execution time of the parallel algorithms when solving problems from the GKLS 4d Hard class}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
   \frametitle{Conclusions}

### 2019-04-07 2019/NUMTA_Barkalov/LNCS/numta19.tex
initial commit

--- 2018-12-28 2019/MOTOR/motor.tex
@@
 \usepackage{graphicx}
+%\usepackage{hyperref}
 % Used for displaying a sample figure. If possible, figure files should
@@
 %
-\title{ Generalized nested optimization scheme for global optimization problems
-\thanks{This study was supported by the Russian Science Foundation, project 
-No.\,16-11-10150.}}
+\title{Acceleration of Global Search through Dual Lipschitz Constant Estimates
+\thanks{This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.}}
 %
@@
 %
-\author{Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
-Ilia Lebedev\orcidID{0000-0002-8736-0652}}
+\author{Roman Strongin%\orcidID{0000-0003-0390-6695} 
+\and Konstantin Barkalov%\orcidID{0000-0001-5273-2471} 
+\and Semen Bevzuk%\orcidID{0000-0002-3845-5356}
+}
 %
-\authorrunning{K. Barkalov, I. Lebedev}
+\authorrunning{R. Strongin et al.}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
-\email{konstantin.barkalov@itmm.unn.ru}}
+\institute{
+Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
+\email{konstantin.barkalov@itmm.unn.ru}
+}
 %
@@
 
-\keywords{Dimensionality reduction \and Global optimization \and Multiextremal functions.}
+\keywords{First keyword  \and Second keyword \and Another keyword.}
 \end{abstract}
@@
 %
-\section{First Section}
-\subsection{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is not indented. The first paragraph that follows a table, figure,
-equation etc. does not need an indent, either.
+\section{Introduction}
 
-Subsequent paragraphs, however, are indented.
+The paper considers global optimization problems with a black-box objective function satisfying the Lipschitz condition. 
+Efficient algorithms for this class of problems require reliable estimates of the Lipschitz constant to be introduced. 
+In fundamental work \cite{Strongin2000} various approaches have been proposed to take into account both global and local properties of the objective function. 
+%В частности, показали свою перспективность  алгоритмы с использованием локальных оценок константы Липшица 
+In particular, algorithms using local estimates of the Lipschitz constant have shown their potential \cite{Sergeyev2010,Sergeyev2016}.
 
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
+The new approach presented in this paper is based on simultaneous use of two estimates: one is substantially larger then the other. 
+The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
+Results of numerical experiments on the random sample of multidimensional functions demonstrate the efficiency of the suggested approach.
 
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
+\section{Conclusion}
 
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
-
-%\begin{figure}
-%\includegraphics[width=\textwidth]{fig1.eps}
-%\caption{A figure caption is always placed below the illustration.
-%Please note that short captions are centered, while long ones are
-%justified by the macro package automatically.} \label{fig1}
-%\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{globalizerSystem}, a
-book~\cite{strSergGO}, proceedings without editors~\cite{Jones2009}.
 %
@@
 % References will then be sorted and formatted in the correct style.
-%
+
 \bibliographystyle{splncs04}
 \bibliography{bibliography}
-%
 
 %\begin{thebibliography}{8}
-%\bibitem{ref_article1}
-%Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-%
-%\bibitem{ref_lncs1}
-%Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-%F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-%Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-%
-%\bibitem{ref_book1}
-%Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-%Location (1999)
-%
-%\bibitem{ref_proc1}
-%Author, A.-B.: Contribution title. In: 9th International Proceedings
-%on Proceedings, pp. 1--2. Publisher, Location (2010)
-%
-%\bibitem{ref_url1}
-%LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-%Oct 2017
 %\end{thebibliography}

### 2019-04-09 2019/RSD_Integer/RSD2019.tex
after translation

@@
 %Russian
-Предложенный алгоритм допускает эффективное использование graphics accelerators. 
-Проведены вычислительные эксперименты при решении серии из 100 multiextremal mixed-integer problems, которые подтверждают хорошее ускорение алгоритма с использованием GPU.
+The proposed algorithm allows an efficient parallelization including the employment of the 
+graphics accelerators. 
+The results of performed numerical experiments (solving a series of 100 multiextremal mixed-
+integer problems) confirm a good speed-up of the algorithm with the use of GPU.
 
@@
 of the whole search domain. 
-As a result, the search of the global optimum is reduced to the construction of a coverage (in general, a nonuniform one) of the space of parameters. 
+As a result, the search of the global optimum is reduced to the construction of a coverage (in 
+general, a nonuniform one) of the space of parameters. 
 The problems, in which some parameters can take the integer values only (mixed-integer global 
@@
 
-The situation when some parameters are featured by the discreteness or integerness is frequent in applied problems.
-As a rule, the integer parameters take a small number of values and may denote, for example, the trademarks of the materials used, the variant of typical layouts of components, etc.
+The situation when some parameters are featured by the discreteness or integerness is frequent 
+in applied problems. As a rule, the integer parameters take a small number of values and may 
+denote, for example, the trademarks of the materials used, the variant of typical layouts of 
+components, etc.
 
@@
 methods of solving the problems of this class are based, as a rule, on the Branch-and-Bound 
-approach \cite{Belotti} or on the Branch-and-Reduce one \cite{Vigerske}. Also, a number of the metaheuristic and genetic 
-algorithms are known, which are based one way or another on the random search concept \cite{Deep,Schluter}.
+approach \cite{Belotti} or on the Branch-and-Reduce one \cite{Vigerske}. Also, a number of 
+the metaheuristic and genetic 
+algorithms are known, which are based one way or another on the random search concept 
+\cite{Deep,Schluter}.
 
@@
 %Russian
-В рамках данного подхода 
+Within the framework of this approach: 
 \begin{itemize}
 	\item 
-	решение многомерных задач сводится  к решению эквивалентных им одномерных, cоответствующая редукция основана на использовании space-filling curves;
+	the solving of the multidimensional problems is reduced to solving the equivalent one-
+dimensional problems; the corresponding reduction is based on the use of the space-filling 
+curves;
 	\item 
-	при решении constrained optimization problems каждое ограничение учитывается и обрабатывается отдельно, штрафные функции не используются;
+	when solving the constrained optimization problems, each constraint is taken into 
+account and processed separately, the penalty functions are not used;
 	\item 
-	распараллеливание процесса поиска осуществляется за счет одновременного вычисления на каждой итерации нескольких значений целевой функции в разных точках search domain.	
+	the parallelization of the search process is performed by means of the simultaneous 
+computing of several objective function values at different points of the search domain within 
+every iteration.
 \end{itemize}
 
-Текст статьи, отражающей результаты проведенного исследования, построен следующим образом. 
-В разделе 2 дано краткое описание описание схемы редукции размерности с использованием space-filling curves, а также описана индексная схема учета ограничений.
-Здесь же приведена формулировка параллельного индексного алгоритма для решения задач непрерывной глобальной оптимизации.
-В разделе 3 предложен подход к обобщению параллельного индексного алгоритма с целью решения mixed-integer problems.
-Приведен способ, позволяющий свести решение mixed-integer problem к решению набора задач непрерывной оптимизации, которое может выполняться параллельно.
-Раздел 4 содержит результаты численных экспериментов.
-Здесь производится сравнение последовательной версии алгоритма с известными аналогами, а также показана эффективность параллельных CPU- и GPU-версий алгоритма при решении серии multiextremal mixed-integer problem.
+The paper text reflecting the results of the performed study is organized in the following way. 
+In Section 2, a brief description of the dimensionality reduction scheme using the space-filling 
+curves is given. Also, the index scheme of accounting for the constraints is described. Here the 
+formulation of the parallel index algorithm for solving the continuous global optimization 
+problems is given as well.
+In Section 3, the approach to the generalization of the parallel index algorithm with the purpose 
+of solving the mixed-integer problems is presented. A method allowing reducing the solving of 
+mixed-integer problem to solving a set of the continuous optimization problems, which can be 
+performed in parallel is given. 
+Section 4 contains the results of numerical experiments. The comparison of the sequential 
+version of the algorithm with the known analogs is conducted here. Also, the efficiency of the 
+parallel CPU- and GPU-versions of the algorithm for solving a series of the multiextremal 
+mixed-integer problem is demonstrated. 
 Section 5 concludes the paper.
@@
 
-By employing the continuous single-valued Peano-Hilbert curve $y(x)$ mapping the unit interval 
+By employing the continuous single-valued Peano-Hilbert curve $y(x)$ mapping the unit 
+interval 
 $[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}), it is possible to find the 
@@
 considered in \cite{Strongin2013}. These evolvents are fractals generated by an iterative
-process, that fill in the hypercube $D$ with accuracy $2^{-m}$, where integer $m>0$ is the evolvent construction parameter.
+process, that fill in the hypercube $D$ with accuracy $2^{-m}$, where integer $m>0$ is the 
+evolvent construction parameter.
 
-Problems of numerical construction of Peano-type space filling curves and the corresponding theory are considered in detail in \cite{Strongin2000,Strongin2013}.
-Examples of the evolvent with different $m$ in two dimensions are given in Fig.~\ref{evolvents}.
+Problems of numerical construction of Peano-type space filling curves and the corresponding 
+theory are considered in detail in \cite{Strongin2000,Strongin2013}. Examples of the evolvent 
+with different $m$ in two dimensions are given in Fig.~\ref{evolvents}.
 
@@
 
-An efficient parallel index algorithm (PIA) for solving the constrained global optimization problem (\ref{problem}) has been developed at University of Nizhni Novgorod.
-The scheme of the algorithm is as follows.
+An efficient parallel index algorithm (PIA) for solving the constrained global optimization 
+problem (\ref{problem}) has been developed at University of Nizhnii Novgorod. The scheme of 
+the algorithm is as follows.
 
-Suppose we have $p \geq 1$  computational elements (e.g., accelerator cores), which can be used to run $p$ trials simultaneously.
-In the first iteration of the method, $p$ trials are run in parallel at various random points $x^i\in(0,1)$, $1\leq i \leq p$.
-Suppose $n \geq 1$  iterations of the method have been completed, and as a result of which, trials were carried out in $k=k(n)$ points $x^i, 1\leq i \leq k$.
-Then the points $x^{k+1},...,x^{k+p}$  of the search trials in the next $(n+1)$-th iteration will be determined according to the rules below.
+Suppose we have $p \geq 1$  computational elements (e.g., accelerator cores), which can be 
+used to run $p$ trials simultaneously. In the first iteration of the method, $p$ trials are run in 
+parallel at various random points $x^i\in(0,1)$, $1\leq i \leq p$. 
+Suppose $n \geq 1$  iterations of the method have been completed, and as a result of which, 
+trials were carried out in $k=k(n)$ points $x^i, 1\leq i \leq k$. Then the points 
+$x^{k+1},...,x^{k+p}$  of the search trials in the next $(n+1)$-th iteration will be determined 
+according to the rules below.
 
@@
 \item 
-Renumber the points $x^1,...,x^k$ from previous iterations with lower indices, lowest to highest coordinate values, i.e.
+Renumber the points $x^1,...,x^k$ from previous iterations with lower indices, lowest to highest 
+coordinate values, i.e.
 \begin{equation}\label{Eq:17}
@@
 \end{equation}
-and match them with the values $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, from (\ref{trial_result}), calculated at these points; points $x_0=0$ и $x_{k+1}=1$ are introduced additionally; the values $z_0$ и $z_{k+1}$ are indeterminate.
+and match them with the values $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, from 
+(\ref{trial_result}), calculated at these points; points $x_0=0$ and $x_{k+1}=1$ are introduced 
+additionally; the values $z_0$ and $z_{k+1}$ are indeterminate.
 \item
-Classify the numbers $i,1\leq i \leq k$, of the trial points from (\ref{Eq:17}) by the number of problem constraints fulfilled at these points, by building the sets
+Classify the numbers $i,1\leq i \leq k$, of the trial points from (\ref{Eq:17}) by the number of 
+problem constraints fulfilled at these points, by building the sets
 \begin{equation}\label{Eq:18}
@@
 \end{equation}
-containing the numbers of all points $x_i,1\leq i \leq k$, with the same values of $\nu$.
-The end points $x_0=0$ and $x_{k+1}=1$ are interpreted as those with zero indices, and they are matched to an additional set $I_0={0,k+1}$.
+containing the numbers of all points $x_i,1\leq i \leq k$, with the same values of $\nu$. The end 
+points $x_0=0$ and $x_{k+1}=1$ are interpreted as those with zero indices, and they are 
+matched to an additional set $I_0={0,k+1}$.
 
@@
 \begin{equation}\label{Eq:20}
-\mu_\nu = \max \left\{ \frac{\left|z_i-z_j\right|}{\left(x_i-x_j\right)^{1/N}} : i,j \in I_\nu, j<i\right\}.
+\mu_\nu = \max \left\{ \frac{\left|z_i-z_j\right|}{\left(x_i-x_j\right)^{1/N}} : i,j \in I_\nu, 
+j<i\right\}.
 \end{equation}
-If the set $I_\nu$ contains less than two elements or $\mu_\nu$ from (\ref{Eq:20}) equals zero, then assume $\mu_\nu=1$.
+If the set $I_\nu$ contains less than two elements or $\mu_\nu$ from (\ref{Eq:20}) equals zero, 
+then assume $\mu_\nu=1$.
 \item
@@
 \end{equation}
-with positive coordinates is called the \textit{reserve vector} and is used as a parameter in the algorithm.
+with positive coordinates is called the \textit{reserve vector} and is used as a parameter in the 
+algorithm.
 \item
@@
 \[
-R(i)=\Delta_i+ \frac{(z_i-z_{i-1})^2}{(r_\nu\mu_\nu)^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu=\nu(x_{i-1})=\nu(x_i),
+R(i)=\Delta_i+ \frac{(z_i-z_{i-1})^2}{(r_\nu\mu_\nu)^2\Delta_i}-2\frac{z_i+z_{i-1}-
+2z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu=\nu(x_{i-1})=\nu(x_i),
 \]
@@
 \]
-where $\Delta_i=(x_i-x_{i-1})^{1/N}$, and the values $r_\nu>1, 1\leq\nu\leq m+1$, are used as parameters in the algorithm.
+where $\Delta_i=(x_i-x_{i-1})^{1/N}$, and the values $r_\nu>1, 1\leq\nu\leq m+1$, are used as 
+parameters in the algorithm.
 \item
@@
 \item
-Carry out $p$ new trials in parallel at the points $x^{k+j}, 1 \leq j \leq p$, calculated by the formulae
+Carry out $p$ new trials in parallel at the points $x^{k+j}, 1 \leq j \leq p$, calculated by the 
+formulae
 \begin{eqnarray*}
 & x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}, \; \nu(x_{t_j-1})\neq \nu(x_{t_j}), \\
-& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}- \frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \; \nu(x_{t_j-1})=\nu(x_{t_j})=\nu. \\
+& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}- \frac{\mathrm{sign}(z_{t_j}-z_{t_j-
+1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \; \nu(x_{t_j-
+1})=\nu(x_{t_j})=\nu. \\
 \end{eqnarray*}
@@
 
-The algorithm stops if the condition $\Delta_{t_j}\leq \epsilon$ becomes true for at least one number $t_j, 1\leq j \leq p$; here  $\epsilon>0$ has an order of magnitude of the desired coordinate accuracy.
+The algorithm stops if the condition $\Delta_{t_j}\leq \epsilon$ becomes true for at least one 
+number $t_j, 1\leq j \leq p$; here  $\epsilon>0$ has an order of magnitude of the desired 
+coordinate accuracy.
 A detailed description of the algorithm convergence theory is presented in \cite{Strongin2000}.
@@
 %Russian
-Рассмотрим теперь случай, когда аргумент функций задачи содержит две компоненты: вектор $y$, принадлежащий гиперинтервалу $D$, и вектор $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е.
+Now let us consider the case when the argument of the problem functions consists of two 
+components: the vector $y$ falling into the hyperinterval $D$ and the vector $u$ taking a finite 
+(and not too large) set of possible values i. e. 
 \begin{gather}\label{problem_i}
@@
 
-Такие конечные наборы могут характеризовать, например, вариант материала, из которого создается объект, геометрические размеры или другие величины, которые могут принадлежать стандартному дискретному ряду, и т.п.
+Such finite sets can characterize, for example, the variant of material, which the object is made 
+from, the geometric sizes, or other quantities, which can belong to a standard discrete series, etc.
 
-Занумеруем целыми числами $s, 1\leq s \leq S,$ все возможные значения вектора $u$, т.е. сопоставим каждому рассматриваемому значению $s$ вектор $u_s$. 
-Тогда рассматриваемая задача может быть записана в виде 
+Let us number by the integer values $s, 1\leq s \leq S,$ all possible values of the vector $u$ i. e. 
+superimpose to each considered value $s$ to the vector $u_s$. 
+Then, the considered problem can be written in the form 
 \begin{gather}\label{problem_is}
- \min_{s\in\{1,...,S\}}\left(\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_i(y,u_s)\leq 0, \; 1 \leq i \leq m\right\}}\right),\\
+ \min_{s\in\{1,...,S\}}\left(\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_i(y,u_s)\leq 0, \; 1 \leq i \leq 
+m\right\}}\right),\\
 D=\left\{ a_j\leq y_j \leq b_j, \; 1 \leq j\leq N \right\}.\nonumber 
@@
 
-Используя схему редукции размерности с помощью развертки $y(x), x\in [0,1]$ можно сопоставить каждой вложенной задаче минимизации по $y$ одномерную задачу минимизации
+Using the dimensionality reduction scheme with the evolvent $y(x), x\in [0,1]$, one can 
+superimpose each nested minimization problem with respect to $y$ to a one-dimensional 
+problem 
 \[
- \min{\left\{ g_{m+1}(y(x),u_s):x \in [0,1], \; g_i(y(x),u_s)\leq 0, \; 1 \leq i \leq m\right\}}, s\in\{1,...,S\}.
+ \min{\left\{ g_{m+1}(y(x),u_s):x \in [0,1], \; g_i(y(x),u_s)\leq 0, \; 1 \leq i \leq m\right\}}, 
+s\in\{1,...,S\}.
 \]
 
-Рассмотрим теперь соответствие 
+Now let us consider the relation 
 \[
@@
 \]
-отображающее любую точку интервала [0,S] на область $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функции 
+mapping any point of the interval [0,S] onto the domain $D$ (the notation $E(x)$ corresponds 
+to the integer part of the number $x$) and define the functions 
 \[
@@
 \]
-имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_k = i, 1\leq i \leq S-1$.
-Значение  $z_k = g_\nu(y(x_k))$ из (\ref{trial_result}) в этих точках будем считать неопределенным, а значения индекса -- равным 0, т.е. $\nu(x_k) = 0$ .
+having, in general, the discontinuities of the first kind at the integer points $x_k = i, 1\leq i \leq 
+S-1$.
+The values $z_k = g_\nu(y(x_k))$ from (\ref{trial_result}) at these points will be considered to 
+be undefined, and the values of the indices -- to equal to 0 i. e. $\nu(x_k) = 0$.
 
-Используя введенные обозначения можно переформулировать исходную задачу как
+Using the introduced notations, one can reformulate the original problem as
 \begin{equation}\label{problem_is1}
@@
 Applying the parallel index algorithm to solving the problem (\ref{problem_is1}), we will find 
-the solution of the problem (\ref{problem_i}). In this case the major part of trials will conducted in 
+the solution of the problem (\ref{problem_i}). In this case the major part of trials will conducted 
+in 
 the subproblem, the solving of which corresponds to the solving of the initial 
-problem (\ref{problem_i}). In the rest subproblems, a only minor part of trials will be performed 
+problem (\ref{problem_i}). In the rest subproblems, only the minor part of trials will be 
+performed 
 since the solutions of these subproblems are the locally optimal ones.
@@
 
-Thus, we have constructed the \textit{Mixed-Integer Parallel Index Algorithm} (MIPIA) based on the  
-reduction of the mixed-integer non-convex optimization problem to the non-convex optimization problem. 
+Thus, we have constructed the \textit{Mixed-Integer Parallel Index Algorithm} (MIPIA) based 
+on the  
+reduction of the mixed-integer non-convex optimization problem to the non-convex 
+optimization problem. 
 %Russian
-Предложенная схема основана на параллельном индексном методе для решения задач непрерывной оптимизации и не является ориентированной на конкретное вычислительное устройство.
-Одной из ключевых операций здесь является параллельное проведение нескольких испытаний в различных точках области поиска (см.
-шаг 7 алгоритма), которе может быть реализовано как на CPU (с использованием OpenMP и/или MPI), так и на GPU (с использованием CUDA).
-Вопросы использования GPU для этих целей были детально рассмотрены в \cite{Barkalov2016,Gergel2016}, указанный подход к использованию GPU был применен и при реализации алгоритма MIPIA.
+The proposed computational scheme is based on the parallel index method for solving the 
+continuous optimization problems and is not oriented onto any particular computational device. 
+Here the parallel performing of several trials at different points of the search domain is one of 
+key operations (see Step 7 of the algorithm), which can be implemented on CPU (with the use 
+of OpenMP and/or MPI) as well as on GPU (with the use CUDA). The issues of the use of 
+GPU for these purposes have been considered in details in 
+Ref.~\cite{Barkalov2016,Gergel2016}. This approach to the use of GPU has been applied in 
+the implementation of the MIPIA algorithm as well.
 
@@
 %Russian
-Первая серия экспериментов была проведена с использованием последовательной версии предложенного алгоритма с целью его сравнения с известными методами аналогичного назначения.
+The first series of experiments was conducted using the sequential version of the proposed 
+algorithm in order to compare this one with well known methods of similar purpose.
 
-Let us compare proposed MIPIA with a genetic algorithm for solving the mixed-integer global optimization problems implemented in Matlab Global Optimization Toolbox \cite{Matlab}. In Table 
+Let us compare proposed MIPIA with a genetic algorithm for solving the mixed-integer global 
+optimization problems implemented in Matlab Global Optimization Toolbox \cite{Matlab}. In 
+Table 
 \ref{tab:1}, the numbers of trials required for solving the known test mixed-integer 
-problems by these methods are presented. For both methods, the same accuracy of search $10^{-2}$ were used. These numerical experiments were conducted on a 
-computer with Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM under MS Windows 10. The 
+problems by these methods are presented. For both methods, the same accuracy of search 
+$10^{-2}$ were used. These numerical experiments were conducted on a 
+computer with Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM under MS Windows 10. 
+The 
 results of experiments have demonstrated the advantage of MIPIA in the number 
@@
 		\hline\noalign{\smallskip}
-	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c}{ GA } & & \multicolumn{2}{c}{MIGSA} \\
+	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c}{ GA } & & 
+\multicolumn{2}{c}{MIGSA} \\
 		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
@@
 
-Последующие серии экспериментов были выполнены для оценки ускорения параллельной версии предложенного алгоритма MIPIA с использованием как CPU, так и GPU.
-В данных экспериментах решалась серия из 100 тестовых mixed-integer problems, сгенерированных случайных образом.
+The next series of experiments were performed in order to evaluate the speedup of the parallel 
+version of the proposed MIPIA algorithm with the use of CPU as well as GPU. In these 
+experiments, a series of 100 mixed-integer test problems generated in a random way was solved.
 
-Известным генератором тестовых задач непрерывной многоэкстремальной оптимизации является GKLS-генератор \cite{Gaviano}.
-Он позволяет генерировать functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.).
-This generator of multiextremal functions is often used for the investigations of the global optimization algorithms~ 
-\cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}. 
-In Fig.~\ref{example} (a) and (b), the contour plots of two-dimensional GKLS functions are presented.
-Figures also shows the points of the trials performed by the method until the required accuracy $\epsilon=10^{-2}$ was achieved.
+GKLS is a well known generator of the test problems for the continuous multiextremal 
+optimization~\cite{Gaviano}. It allows generating the functions of arbitrary dimensionality with 
+known properties (the number of local minima, the size of their domains of attraction, the global 
+minimizer, etc.).  
+This generator of multiextremal functions is used for the investigations of the global 
+optimization algorithms 
+often~\cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}. 
+In Fig.~\ref{example} (a) and (b), the contour plots of two-dimensional GKLS functions are 
+presented. Figures also shows the points of the trials performed by the method until the required 
+accuracy $\epsilon=10^{-2}$ was achieved.
 
@@
 
-В проведенных экспериментах GKLS-генератор использовался как основа для построения mixed-integer problems. Правила, позволяющие генерировать тестовые задачи подобного вида, заключаются в следующем.
+In the performed experiments, the GKLS generator was used as a base for the construction the 
+mixed-integer problems. The rules, allowing generating the test problems of this kind consist in 
+the following.
 
 \begin{enumerate}
-	\item С помощью генератора GKLS порождается непрерывная многоэкстремальная функция $\varphi(y), \; y\in D, \; D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\}$.
-Глобальный минимум данной функции достигается в известной точке $y'=(y'_1,...,y'_N)$ и равен $\varphi'=\varphi(y')=-1$.
-	\item Генерируется concave mixed-integer function 
+	\item A continuous multiextremal function $\varphi(y), \; y\in D, \; D = \left\{ a_j\leq 
+y_j\leq b_j, 1\leq j \leq N \right\}$ is generated with the use of the GKLS generator. The global 
+minimum of this function is achieved at the known point $y'=(y'_1,...,y'_N)$ and equals 
+$\varphi'=\varphi(y')=-1$.
+	\item A concave mixed-integer function 
 	\[
-			\mu(y,u) = -2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} \right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right],
+			\mu(y,u) = -2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} 
+\right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right],
 	\]
-	где 
+	is generated where 
 	\begin{gather}
@@
 	\end{gather}
-	т.е. данная функция имеет $N$ непрерывных и $M$ дискретных параметров и достигает минимального значения в точке $(y',b)$.
-	\item Вычисляется коэффициент 
+	i. e. this function has $N$ continuous and $M$ discrete parameters and achieves its 
+minimum value at the point $(y',b)$.
+	\item The coefficient 
 	\[
@@
 	\]
-	Очевидно, что минимум вогнутой функции $\mu(y,u)$ в одной из угловых точек области поиска, поэтому при размерности задачи порядка 10 его вычисление может быть выполнено методом перебора.
-	\item Формируется многоэкстремальная mixed-integer function 
+	is computed. 
+It is obvious, that the minimum of a concave function $\mu(y,u)$ is located at one of the corner 
+points of the search domain. Therefore, if the problem dimensionality is of the order of 10 it can 
+be found by the brute force method.
+	\item A multiextremal mixed-integer function 
 	\[
@@
 	\]
-	По построению $f(y,u)$  будет принимать минимальное значение в точке $(y',b)$.
+	is formed.
+By construction, $f(y,u)$ would take its minimum value at the point $(y',b)$.
 
@@
 
-In the problems generated in our experiments, there were 5 discrete and 6 continuous parameters 
+In the problems generated in our experiments, there were 5 discrete and 6 continuous 
+parameters 
 	\begin{gather}
@@
 
-Было сгенерировано 100 11-и мерных mixed-integer problems данного типа. The accuracy of the search was equal to $10^{-2}$.   
+Total 100 11-dimensional mixed-integer problems of this kind were generated. The accuracy of 
+the search was equal to $10^{-2}$.   
 Computational experiments were carried out on Lobachevsky supercomputer. The node of 
 supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The 
-CPU had 8 cores, i.e. each node had a total of 16 cores and two NVIDIA Kepler K20Х GPUs.
+CPU had 8 cores, i.e. each node had a total of 16 cores and two NVIDIA Kepler K20X GPUs.
 
-В Table \ref{tab:2} приведены результаты экспериментов на CPU с использованием OpenMP в зависимости от числа задействованных threads $p$.
-Использовалось 1, 8 и 16 потоков.
-Приведено среднее число итераций $K_{av}$, требуемое для решения задачи, среднее время решения $T_{av}$ (в секундах), time speedup $S$ and iteration speedup $s$ (по отношению к последовательному запуску, т.е.
-$p=1$).
-В соответствии со схемой распараллеливания число испытаний на одной итерации параллельного алгоритма равнялось числу задействованных threads.
+In Table \ref{tab:2}, the results of experiments on CPU with the use of OpenMP are presented 
+subject to the number of employed threads $p$. Total 1, 8, and 16 threads were used. The 
+average number of iterations $K_{av}$ required to solve the problem, the average time of 
+solving $T_{av}$ (in seconds), the time speedup $S$, and iteration speedup $s$ (with respect 
+to the sequential run i. e. for $p=1$) are presented. In accordance with the parallelization 
+scheme, the number of trials within a single iteration of the parallel algorithm was equal to the 
+number of employed threads.
 
 \begin{table}
-	\caption{Результаты экспериментов на CPU}
+	\caption{The results of experiments on CPU}
 	\label{tab:2}
@@
 
-Как видно из Table \ref{tab:2} , наблюдается почти линейное ускорение по итерациям и значительное ускорение по времени решения задачи, при этом последовательному алгоритму в среднем требовалось примерно 1.5 часа для решения задачи.
+As one can see from Table \ref{tab:2}, almost linear speedup in iterations and an essential 
+speedup in the problem solving time have been observed. At that, the sequential algorithm spent 
+almost 1.5 hours in average to solve a problem.
 
-В Table \ref{tab:3} приведены результаты экспериментов на GPU с использованием CUDA в зависимости от числа задействованных GPU-threads.
-Приведено среднее число итераций $K_{av}$, требуемое для решения задачи, среднее время решения $T_{av}$ (в секундах), time speedup $S$ and iteration speedup $s$ (по отношению к полной загрузке CPU на узле кластера).
+In Table \ref{tab:3}, the results of experiments on GPU with the use of CUDA are presented 
+subject to the number of GPU threads employed. The average number of iterations $K_{av}$ 
+required to solve a problem, the average solving time $T_{av}$ (in seconds), the time speedup 
+$S$, and iteration speedup $s$ (with respect to the full load of CPU on a cluster node) are 
+presented.
 
 \begin{table}
-	\caption{Результаты экспериментов на GPU}
+	\caption{The results of experiments on GPU}
 	\label{tab:3}
@@
 
-Результаты эксперимнетов показывают почти одинаковое time speedup и линейное iteration speedup при использовании $p=256, 512, 1024$ threads.
-Однако при использовании $p=2048$ threads время работы алгоритма замедляется, но число итераций продолжает падать.
-Данный эффект объясняется тем, что для параллельного проведения $p$ испытаний используется GPU, а для обработки результатов испытаний (которая подразумевает работу со всей накопленной на предыдущих итерациях поисковой информацией) используется CPU.
-И при использовании $p=2048$ GPU threads время, которое тратится на передачу и обработку результатов 2048 испытаний, становится сопоставимым со временем проведения испытания, что и приводит к замедлению алгоритма в целом.
+The results of experiments demonstrate almost the same time speedup and linear iteration 
+speedup when using $p=256, 512,$ and $1024$ threads. However, when using $p=2048$ 
+threads, the algorithm worktime increased but the number of iterations continued to decrease. 
+This effect was explained by the fact that for parallel running of $p$ trials GPU is used but  for 
+processing the results of the trials (which implies the processing of the whole search information 
+accumulated during the preceding iterations) CPU is uses. And when employing $p=2048$ 
+GPU threads, the time, which is spent for the transfer and processing of the results of 2048 
+trials became comparable to the time of executing the trials that leads to the slowing down of 
+the algorithm as a whole.
 
@@
 %Russian
-В данной статье представлены результаты, полученные в Lobachevsky State University of Nizhni Novgorod при разработке и  исследовании параллельных алгоритмов глобальной оптимизации для решения многоэкстремальных задач, в которых часть параметров является непрерывными, а часть - дискретными.
-Для задач указанного класса был предложен эффективный параллельный алгоритм их решения алгоритм. 
-В последовательном варианте данный алгоритм не уступает алгоритму аналогичного назначения, реализованному в Matlab Global Optimization Toolbox.
-В параллельном варианте предложенный алгоритм допускает эффективную реализацию как на CPU, так и на GPU. 
-Проведены вычислительные эксперименты на серии из 100 multiextremal mixed-integer problems, убедительно демонстрирующие хорошее ускорение алгоритма с использованием GPU.
-Так, параллельный алгоритм с использованием GPU показал time speedup $S = 182$ по отношению к последовательному алгоритму и $S = 23.7$ по отношению к алгоритму, полностью задействующему два CPU на узле кластера.
+In the present paper, the results obtained in Lobachevskii State University of Nizhnii Novgorod 
+when developing and investigating the parallel global optimization algorithms for solving the 
+multiextremal problems, in which some parameters are continuous and the others are the discrete 
+ones are presented. 
+An efficient parallel algorithm has been proposed for solving the problems of the specified class. 
+In the sequential variant, this algorithm does not inferior to similar purpose algorithm 
+implemented in Matlab Global Optimization Toolbox.
+In the parallel variant, the proposed algorithm allows the efficient implementations on CPU as 
+well as on GPU. 
+The numerical experiments
+on a series of 100 multiextremal mixed-integer test problems have been carried out 
+demonstrating a good speedup of the algorithm with the use of GPU convincingly.
+Thus, the parallel algorithm with the use of GPU demonstrated the time speedup $S = 182$ 
+relative to the sequential one and $S = 23.7$ relative to the algorithm employing two CPU on 
+the cluster node fully.
 
@@
 \bibitem{Burer}
-Burer, S., Letchford, A.N.: Non-convex mixed-integer nonlinear programming: A survey. Surveys in Operations Research and Management Science \textbf{17}, 97--106 (2012)
+Burer, S., Letchford, A.N.: Non-convex mixed-integer nonlinear programming: A survey. 
+Surveys in Operations Research and Management Science \textbf{17}, 97--106 (2012)
 
 \bibitem{Boukouvala}
-Boukouvala, F., Misener, R., Floudas, C.A.: Global optimization advances in Mixed-Integer Nonlinear Programming, MINLP, and Constrained Derivative-Free Optimization, CDFO.
-European J.
-Oper.
-Res.
-\textbf{252}, 701--727 (2016)
+Boukouvala, F., Misener, R., Floudas, C.A.: Global optimization advances in Mixed-Integer 
+Nonlinear Programming, MINLP, and Constrained Derivative-Free Optimization, CDFO. 
+European J. Oper. Res. \textbf{252}, 701--727 (2016)
 
 \bibitem{Deep}
-Deep, K., Singh, K. P., Kansal, M.L., Mohan, C.: A real coded genetic algorithm for solving integer and mixed integer optimization problems. Appl. Math. Comput. \textbf{212}(2), 505--518 (2009)
+Deep, K., Singh, K. P., Kansal, M.L., Mohan, C.: A real coded genetic algorithm for solving 
+integer and mixed integer optimization problems. Appl. Math. Comput. \textbf{212}(2), 505--
+518 (2009)
 
 \bibitem{Schluter}
-Schl\"uter, M., Egea, J.A., Banga, J.R.: Extended ant colony optimization for non-convex mixed integer nonlinear programming. Computers and Operations Research \textbf{36}(7), 2217--2229 (2009)
+Schl\"uter, M., Egea, J.A., Banga, J.R.: Extended ant colony optimization for non-convex 
+mixed integer nonlinear programming. Computers and Operations Research \textbf{36}(7), 
+2217--2229 (2009)
 
 \bibitem{Belotti}
-Belotti, P., Lee, J., Liberti, L., Margot, F., W\"achter, A.: Branching and bounds tightening techniques for non-convex MINLP. Optim. Method. Softw. \textbf{24}(4-5), 597--634 (2009)
+Belotti, P., Lee, J., Liberti, L., Margot, F., W\"achter, A.: Branching and bounds tightening 
+techniques for non-convex MINLP. Optim. Method. Softw. \textbf{24}(4-5), 597--634 (2009)
 
 \bibitem{Vigerske}
-Vigerske, S., Gleixner, A.: SCIP: global optimization of mixed-integer nonlinear programs in a branch-and-cut framework. 
+Vigerske, S., Gleixner, A.: SCIP: global optimization of mixed-integer nonlinear programs in a 
+branch-and-cut framework. 
 Optim. Method. Softw. \textbf{33}(3), 563--593 (2018)
@@
 \bibitem{Strongin2000}
-Strongin, R.G., Sergeyev, Y.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000) %; DOI: 10.1007/978-1-4615-4677-1
+Strongin, R.G., Sergeyev, Y.D.: Global optimization with non-convex constraints. Sequential 
+and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000) %; DOI: 10.1007/978-
+1-4615-4677-1
 
 \bibitem{Strongin2013}
-Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013) %;  DOI: 10.1007/978-1-4614-8042-6
+Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-
+filling curves. Springer (2013) %;  DOI: 10.1007/978-1-4614-8042-6
 
 \bibitem{Floudas}
-Floudas, C.A., Pardalos, P.M.:  Handbook of test problems in local and global optimization. Springer (1999)  %; DOI: 10.1007/978-1-4757-3040-1
+Floudas, C.A., Pardalos, P.M.:  Handbook of test problems in local and global optimization. 
+Springer (1999)  %; DOI: 10.1007/978-1-4757-3040-1
 
@@
 
-\bibitem{Gaviano} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
-ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003)
+\bibitem{Gaviano} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for 
+generation of classes of test functions with known local and global minima for global 
+optimization. ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003)
 
 \bibitem{Paulavicius2014} 
-Paulavi\v{c}ius, R., Sergeyev, Y., Kvasov, D., \v{Z}ilinskas, J.: Globally-biased DISIMPL algorithm for expensive global optimization. J. Glob. Optim. \textbf{59}(2-3), 545--567 (2014)
+Paulavi\v{c}ius, R., Sergeyev, Y., Kvasov, D., \v{Z}ilinskas, J.: Globally-biased DISIMPL 
+algorithm for expensive global optimization. J. Glob. Optim. \textbf{59}(2-3), 545--567 (2014)
 
 \bibitem{SergeyevKvasov2015} 
-Sergeyev, Y.D., Kvasov, D.E.: A deterministic global optimization using smooth diagonal auxiliary functions. Commun. Nonlinear. Sci. Numer. Simulat. \textbf{21}(1-3), 99--111 (2015)
+Sergeyev, Y.D., Kvasov, D.E.: A deterministic global optimization using smooth diagonal 
+auxiliary functions. Commun. Nonlinear. Sci. Numer. Simulat. \textbf{21}(1-3), 99--111 (2015)
 
 \bibitem{Lebedev2015}
-Lebedev, I., Gergel, V.: Heterogeneous parallel computations for solving global optimization problems. Procedia Computer Science \textbf{66}, 53--62 (2015)
+Lebedev, I., Gergel, V.: Heterogeneous parallel computations for solving global optimization 
+problems. Procedia Computer Science \textbf{66}, 53--62 (2015)
 
 \bibitem{Gergel2015}
-Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of computationally intensive multiextremal optimization problems.
-Lecture Notes in Computer Science  \textbf{9251}, 505--515 (2015)
+Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of 
+computationally intensive multiextremal optimization problems. Lecture Notes in Computer 
+Science  \textbf{9251}, 505--515 (2015)
 
 \bibitem{Barkalov2016} 
-Barkalov, K., Gergel, V.: Parallel global optimization on GPU. Journal of Global Optimization 66(1), 2--20 (2016)
+Barkalov, K., Gergel, V.: Parallel global optimization on GPU. Journal of Global Optimization 
+66(1), 2--20 (2016)
 
 \bibitem{Gergel2016}
-Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster. In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings, vol. 1738, art. no. 400006 (2016)
+Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster. 
+In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings, vol. 1738, art. no. 400006 
+(2016)
 

### 2019-04-10 2019/RSD_Integer/RSD2019.tex
Before correction

@@
 mixed-integer problem to solving a set of the continuous optimization problems, which can be 
-performed in parallel is given. 
+performed in parallel, is given. 
 Section 4 contains the results of numerical experiments. The comparison of the sequential 
@@
 An efficient parallel index algorithm (PIA) for solving the constrained global optimization 
-problem (\ref{problem}) has been developed at University of Nizhnii Novgorod. The scheme of 
+problem (\ref{problem}) has been developed at University of Nizhni Novgorod. The scheme of 
 the algorithm is as follows.
@@
 Now let us consider the case when the argument of the problem functions consists of two 
-components: the vector $y$ falling into the hyperinterval $D$ and the vector $u$ taking a finite 
-(and not too large) set of possible values i. e. 
+components: the vector $y$ from the hyperinterval $D$ and the vector $u$ taking a finite 
+(and not too large) set of possible values, i.e. 
 \begin{gather}\label{problem_i}
@@
 
-Let us number by the integer values $s, 1\leq s \leq S,$ all possible values of the vector $u$ i. e. 
-superimpose to each considered value $s$ to the vector $u_s$. 
+Let us number by the integer values $s, 1\leq s \leq S,$ all possible values of the vector $u$, i.e. 
+juxtapose each considered value $s$ with the vector $u_s$. 
 Then, the considered problem can be written in the form 
@@
 \]
-having, in general, the discontinuities of the first kind at the integer points $x_k = i, 1\leq i \leq 
+having, in general, the jump discontinuities at the integer points $x_k = i, 1\leq i \leq 
 S-1$.
 The values $z_k = g_\nu(y(x_k))$ from (\ref{trial_result}) at these points will be considered to 
-be undefined, and the values of the indices -- to equal to 0 i. e. $\nu(x_k) = 0$.
+be undefined, and the values of the indices -- to equal to 0, i.e. $\nu(x_k) = 0$.
 
@@
 As an illustration, Fig. \ref{fig:1} presents the plots of the functions corresponding to a 
-problem with one continuous parameter and one binary parameter.
+problem 
+\[
+%\min{\left\{ x^2 - 10\cos(2\pi x) + u^2 - 10\cos(2\pi u) : x\in [-2.22, 1.8], u \in {0,1} \right\}}
+\min{\left\{ u^2 (\sin(x) +\sin(10x/3) : x\in [2.7, 7.5], u \in {1,2} \right\}}
+\]
+with one continuous parameter $x$ and one binary parameter $u$.
+%Еще один график?
 
@@
 
-Thus, we have constructed the \textit{Mixed-Integer Parallel Index Algorithm} (MIPIA) based 
-on the  
-reduction of the mixed-integer non-convex optimization problem to the non-convex 
-optimization problem. 
+Thus, we have constructed the \textit{Mixed-Integer Parallel Index Algorithm} \break 
+(MIPIA) based on the  reduction of the mixed-integer non-convex optimization problem to the non-convex optimization problem. 
 %Russian
@@
 GPU for these purposes have been considered in details in 
-Ref.~\cite{Barkalov2016,Gergel2016}. This approach to the use of GPU has been applied in 
+\cite{Barkalov2016,Gergel2016}. This approach to the use of GPU has been applied in 
 the implementation of the MIPIA algorithm as well.
@@
 optimization problems implemented in Matlab Global Optimization Toolbox \cite{Matlab}. In 
-Table 
-\ref{tab:1}, the numbers of trials required for solving the known test mixed-integer 
+Table \ref{tab:1}, the numbers of trials required for solving the known test mixed-integer 
 problems by these methods are presented. For both methods, the same accuracy of search 
@@
 computer with Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM under MS Windows 10. 
-The 
-results of experiments have demonstrated the advantage of MIPIA in the number 
+The results of experiments have demonstrated the advantage of MIPIA in the number 
 of iterations as well as in the execution time.
@@
 experiments, a series of 100 mixed-integer test problems generated in a random way was solved. 
+Computational experiments were carried out on Lobachevsky supercomputer. The node of 
+supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The 
+CPU had 8 cores, i.e. each node had a total of 16 cores and two NVIDIA Kepler K20X GPUs.
 
-GKLS is a well known generator of the test problems for the continuous multiextremal 
-optimization~\cite{Gaviano}. It allows generating the functions of arbitrary dimensionality with 
+GKLS~\cite{Gaviano} is a well known generator of the test problems for the continuous multiextremal 
+optimization. It allows generating the functions of arbitrary dimensionality with 
 known properties (the number of local minima, the size of their domains of attraction, the global 
 minimizer, etc.).  
-This generator of multiextremal functions is used for the investigations of the global 
-optimization algorithms 
-often~\cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}. 
+This generator of multiextremal functions is often used for the investigations of the global 
+optimization algorithms ~\cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}. 
 In Fig.~\ref{example} (a) and (b), the contour plots of two-dimensional GKLS functions are 
@@
 In the performed experiments, the GKLS generator was used as a base for the construction the 
-mixed-integer problems. The rules, allowing generating the test problems of this kind consist in 
+mixed-integer problems. The rules allowing generating the test problems of this type consist in 
 the following.
@@
 \begin{enumerate}
-	\item A continuous multiextremal function $\varphi(y), \; y\in D, \; D = \left\{ a_j\leq 
+	\item A continuous multiextremal function $\varphi(y), y\in D = \left\{ a_j\leq 
 y_j\leq b_j, 1\leq j \leq N \right\}$ is generated with the use of the GKLS generator. The global 
@@
 	y\in D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\} \subset R^N,\nonumber \\
-	u\in U = \left\{ u_j \in  \left\{a_j, ..., b_j \right\}, 1\leq j \leq M \right\};\nonumber
+	u\in U = \left\{ u_j \in  \left\{a_j, ..., b_j \right\}, 1\leq j \leq M \right\}.\nonumber
 	\end{gather}
-	i. e. this function has $N$ continuous and $M$ discrete parameters and achieves its 
+	This function has $N$ continuous and $M$ discrete parameters and achieves its 
 minimum value at the point $(y',b)$.
@@
 
-Total 100 11-dimensional mixed-integer problems of this kind were generated. The accuracy of 
-the search was equal to $10^{-2}$.   
-Computational experiments were carried out on Lobachevsky supercomputer. The node of 
-supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The 
-CPU had 8 cores, i.e. each node had a total of 16 cores and two NVIDIA Kepler K20X GPUs.
+A hundred 11-dimensional mixed-integer problems of this type were generated in total. 
+For the purpose of simulation of the computational complexity inherent to
+applied problems of optimization, calculation of the objective function in all performed
+experiments was made more complex by additional calculations without changing the type
+of function and arrangement of its minima (series summation from 80 thousand elements).
+The accuracy of the search was equal to $10^{-2}$.%Параметры метода!!!
 
@@
 solving $T_{av}$ (in seconds), the time speedup $S$, and iteration speedup $s$ (with respect 
-to the sequential run i. e. for $p=1$) are presented. In accordance with the parallelization 
+to the sequential run, i.e. for $p=1$) are presented. In accordance with the parallelization 
 scheme, the number of trials within a single iteration of the parallel algorithm was equal to the 
@@
 
-As one can see from Table \ref{tab:2}, almost linear speedup in iterations and an essential 
+As one can see from Table \ref{tab:2}, almost linear speedup in iterations and an significant 
 speedup in the problem solving time have been observed. At that, the sequential algorithm spent 
@@
 %Russian
-In the present paper, the results obtained in Lobachevskii State University of Nizhnii Novgorod 
+In the present paper, the results obtained in Lobachevskii State University of Nizhni Novgorod 
 when developing and investigating the parallel global optimization algorithms for solving the 
@@
 well as on GPU. 
-The numerical experiments
-on a series of 100 multiextremal mixed-integer test problems have been carried out 
-demonstrating a good speedup of the algorithm with the use of GPU convincingly.
+The numerical experiments on a series of 100 multiextremal mixed-integer test problems have been carried out 
+convincingly demonstrating a good speedup of the algorithm with the use of GPU.
 Thus, the parallel algorithm with the use of GPU demonstrated the time speedup $S = 182$ 
-relative to the sequential one and $S = 23.7$ relative to the algorithm employing two CPU on 
-the cluster node fully.
+relative to the sequential one and $S = 23.7$ relative to the algorithm fully employing two CPU on 
+the cluster node.
 
@@
 
+\bibitem{Belotti}
+Belotti, P., Lee, J., Liberti, L., Margot, F., W\"achter, A.: Branching and bounds tightening 
+techniques for non-convex MINLP. Optim. Method. Softw. \textbf{24}(4-5), 597--634 (2009)
+
+\bibitem{Vigerske}
+Vigerske, S., Gleixner, A.: SCIP: global optimization of mixed-integer nonlinear programs in a 
+branch-and-cut framework. 
+Optim. Method. Softw. \textbf{33}(3), 563--593 (2018)
+
 \bibitem{Deep}
@@
 
-\bibitem{Belotti}
-Belotti, P., Lee, J., Liberti, L., Margot, F., W\"achter, A.: Branching and bounds tightening 
-techniques for non-convex MINLP. Optim. Method. Softw. \textbf{24}(4-5), 597--634 (2009)
-
-\bibitem{Vigerske}
-Vigerske, S., Gleixner, A.: SCIP: global optimization of mixed-integer nonlinear programs in a 
-branch-and-cut framework. 
-Optim. Method. Softw. \textbf{33}(3), 563--593 (2018)
-
 \bibitem{Strongin2000}
@@
 
+\bibitem{Barkalov2016} 
+Barkalov, K., Gergel, V.: Parallel global optimization on GPU. Journal of Global Optimization 
+66(1), 2--20 (2016)
+
+\bibitem{Gergel2016}
+Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster. 
+In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings, vol. 1738, art. no. 400006 
+(2016)
+
+\bibitem{Matlab}
+https://www.mathworks.com/help/gads/mixed-integer-optimization.html
+
 \bibitem{Floudas}
@@
 Springer (1999)  %; DOI: 10.1007/978-1-4757-3040-1
-
-\bibitem{Matlab}
-https://www.mathworks.com/help/gads/mixed-integer-optimization.html
 
@@
 
-\bibitem{Barkalov2016} 
-Barkalov, K., Gergel, V.: Parallel global optimization on GPU. Journal of Global Optimization 
-66(1), 2--20 (2016)
-
-\bibitem{Gergel2016}
-Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster. 
-In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings, vol. 1738, art. no. 400006 
-(2016)
-
 \end{thebibliography}

### 2019-04-11 2019/RSD_Integer/RSD2019.tex
Corrected

@@
 
-In the present paper, the mixed-integer global optimization problems are considered. A novel 
-parallel algorithm for solving the problems of this class based on the information-statistical 
-approach to solving the continuous global optimization problems has been proposed. 
-The comparison of this algorithm with known analogs demonstrating the efficiency of the 
-developed approach has been conducted. 
+The paper considers the mixed-integer global optimization problems. A novel 
+parallel algorithm for solving the problems of this class based on the index 
+algorithm to solving the continuous global optimization problems has been proposed. 
+The comparison of this algorithm with known analogs demonstrates the efficiency of the 
+developed approach. 
 %Russian
 The proposed algorithm allows an efficient parallelization including the employment of the 
-graphics accelerators. 
-The results of performed numerical experiments (solving a series of 100 multiextremal mixed-
-integer problems) confirm a good speed-up of the algorithm with the use of GPU.
+graphics accelerators. The results of performed numerical experiments (solving a series of 100 multiextremal mixed-integer problems) confirm a good speedup of the algorithm with the use of GPU.
 
@@
 
-In the present paper, the global optimization problems and the method of solving these ones are 
-considered. The global optimization problems are the computation-costly ones since the global 
+In this paper the global optimization problems and the method of solving these ones are 
+considered. The global optimization problems are the time-consuming ones since the global 
 optimum is an integral characteristic of the problem being solved and requires the investigation 
@@
 general, a nonuniform one) of the space of parameters. 
-The problems, in which some parameters can take the integer values only (mixed-integer global 
-optimization problems) are of special interest because for these ones it is more difficult to build 
-the estimates of the optimum as compared to the continuous problems.
+The problems, in which some parameters can take the discrete or integer values only (mixed-integer global 
+optimization problems) are of special interest because for these problems it is more difficult to build 
+the estimates of the optimum as compared to the continuous ones.
 
@@
 methods of solving the problems of this class are based, as a rule, on the Branch-and-Bound 
-approach \cite{Belotti} or on the Branch-and-Reduce one \cite{Vigerske}. Also, a number of 
+\cite{Belotti} or on the Branch-and-Reduce approach \cite{Vigerske}. Also, a number of 
 the metaheuristic and genetic 
@@
 In the present study, we proposed a novel parallel method for solving the mixed-integer 
-problems based on the information-statistical approach to solving the global optimization 
+problems based on the index approach to solving the constrained global optimization 
 problems \cite{Strongin2000,Strongin2013}. 
@@
 	\item 
-	the solving of the multidimensional problems is reduced to solving the equivalent one-
-dimensional problems; the corresponding reduction is based on the use of the space-filling 
-curves;
+	the solving of the multidimensional problems is reduced to solving the equivalent one-dimensional problems; the corresponding reduction is based on the use of the space-filling curves;
 	\item 
@@
 In Section 3, the approach to the generalization of the parallel index algorithm with the purpose 
-of solving the mixed-integer problems is presented. A method allowing reducing the solving of 
+of solving the mixed-integer problems is presented. A method that allows to reduce solving of 
 mixed-integer problem to solving a set of the continuous optimization problems, which can be 
@@
 \]
-Algorithms for numerical construction of Hilbert curve approximation (\textit{evolvent}) are 
-considered in \cite{Strongin2013}. These evolvents are fractals generated by an iterative
+Algorithms for numerical construction of Peano-Hilbert curve approximation (\textit{evolvent}) are 
+considered in \cite{Strongin2000,Strongin2013}. These evolvents are fractals generated by an iterative
 process, that fill in the hypercube $D$ with accuracy $2^{-m}$, where integer $m>0$ is the 
-evolvent construction parameter.
-
-Problems of numerical construction of Peano-type space filling curves and the corresponding 
-theory are considered in detail in \cite{Strongin2000,Strongin2013}. Examples of the evolvent 
-with different $m$ in two dimensions are given in Fig.~\ref{evolvents}.
+evolvent construction parameter. Examples of the evolvent with different $m$ in two dimensions are given in Fig.~\ref{evolvents}.
 
@@
 
-Suppose we have $p \geq 1$  computational elements (e.g., accelerator cores), which can be 
+Suppose we have $p \geq 1$  computational elements (e.g., processor cores), which can be 
 used to run $p$ trials simultaneously. In the first iteration of the method, $p$ trials are run in 
@@
 (\ref{trial_result}), calculated at these points; points $x_0=0$ and $x_{k+1}=1$ are introduced 
-additionally; the values $z_0$ and $z_{k+1}$ are indeterminate.
+additionally, the values $z_0$ and $z_{k+1}$ are indeterminate.
 \item
@@
 points $x_0=0$ and $x_{k+1}=1$ are interpreted as those with zero indices, and they are 
-matched to an additional set $I_0={0,k+1}$.
+matched to an additional set $I_0= \left\{ 0,k+1 \right\}$.
 
@@
 \end{equation}
-where $M$ is the maximum current value of the index, and the vector
-\begin{equation}\label{Eq:22}
-\epsilon _R=\left(\epsilon_1,...,\epsilon_m\right),
-\end{equation}
-with positive coordinates is called the \textit{reserve vector} and is used as a parameter in the 
+where $M$ is the maximum current value of the index, and the vector $\epsilon _R=\left(\epsilon_1,...,\epsilon_m\right)$ with positive coordinates is called the \textit{reserve vector} and is used as a parameter in the
 algorithm.
@@
 coordinate accuracy.
+
+This method of organizing parallel computing has the following justification.
+The characteristics of intervals $R(i)$ used in the index algorithm can be considered as
+probability measures of the global minimum point location in these intervals. Inequalities
+(\ref{Eq:23}) arrange intervals according to their characteristics, and trials are carried out in parallel in
+the first $p$ intervals with the largest probabilities.
 A detailed description of the algorithm convergence theory is presented in \cite{Strongin2000}.
@@
 \begin{gather}\label{problem_is}
- \min_{s\in\{1,...,S\}}\left(\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_i(y,u_s)\leq 0, \; 1 \leq i \leq 
-m\right\}}\right),\\
+ \min_{s\in\{1,...,S\}} \left\{ \min_{}{\left\{ g_{m+1}(y,u_s):y\in D, \; g_i(y,u_s)\leq 0, \; 1 \leq i \leq 
+m\right\}}\right\},\\
 D=\left\{ a_j\leq y_j \leq b_j, \; 1 \leq j\leq N \right\}.\nonumber 
@@
 %\min{\left\{ x^2 - 10\cos(2\pi x) + u^2 - 10\cos(2\pi u) : x\in [-2.22, 1.8], u \in {0,1} \right\}}
-\min{\left\{ u^2 (\sin(x) +\sin(10x/3) : x\in [2.7, 7.5], u \in {1,2} \right\}}
+\min{\left\{ u^2 (\sin(x) +\sin(10x/3) : x\in [2.7, 7.5], u \in \{1,2\} \right\}}
 \]
@@
 	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c}{ GA } & & 
-\multicolumn{2}{c}{MIGSA} \\
+\multicolumn{2}{c}{MIPIA} \\
 		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
@@
 minimum of this function is achieved at the known point $y'=(y'_1,...,y'_N)$ and equals 
-$\varphi'=\varphi(y')=-1$.
+$\varphi(y')=-1$.
 	\item A concave mixed-integer function 
 	\[
-			\mu(y,u) = -2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} 
+			h(y,u) = -2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} 
 \right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right],
@@
 	\[
-	C = 4 - \min_{y,u} \left\{ \mu(y,u) \right\}, \; y\in D, u \in U.
+	h_{min} = 4 - \min_{y,u} \left\{ h(y,u) \right\}, \; y\in D, u \in U.
 	\]
@@
 	\[
-	f(y,u) = \left(\varphi(y) + \sum_{j=1}^M{u_j}\right)\left(C + \mu(y,u)\right).
+	f(y,u) = \left(\varphi(y) + \sum_{j=1}^M{u_j}\right)\left(h_{min} + h(y,u)\right).
 	\]
@@
 For the purpose of simulation of the computational complexity inherent to
-applied problems of optimization, calculation of the objective function in all performed
+applied optimization problems, calculation of the objective function in all performed
 experiments was made more complex by additional calculations without changing the type
@@
 %Russian
-In the present paper, the results obtained in Lobachevskii State University of Nizhni Novgorod 
+In the paper the results obtained in Lobachevsky State University of Nizhni Novgorod 
 when developing and investigating the parallel global optimization algorithms for solving the 

### 2019-04-11 2019/RSD_Integer/RSD2019.tex
Semi-final

@@
 parallel algorithm for solving the problems of this class based on the index 
-algorithm to solving the continuous global optimization problems has been proposed. 
+algorithm for solving the continuous global optimization problems has been proposed. 
 The comparison of this algorithm with known analogs demonstrates the efficiency of the 
@@
 
-In this paper the global optimization problems and the method of solving these ones are 
+In this paper the global optimization problems and the method of their solving are 
 considered. The global optimization problems are the time-consuming ones since the global 
@@
 general, a nonuniform one) of the space of parameters. 
-The problems, in which some parameters can take the discrete or integer values only (mixed-integer global 
-optimization problems) are of special interest because for these problems it is more difficult to build 
+The problems, in which some parameters can take the discrete or integer values only (\textit{mixed-integer global 
+optimization problems}) are of special interest because for these problems it is more difficult to build 
 the estimates of the optimum as compared to the continuous ones.
@@
 \]
-with one continuous parameter $x$ and one binary parameter $u$.
+with one continuous parameter $x$ and one integer parameter $u$.
 %Еще один график?
@@
 	\[
-	h_{min} = 4 - \min_{y,u} \left\{ h(y,u) \right\}, \; y\in D, u \in U.
+	h_{max} = 4 - \min_{y,u} h(y,u) , \; y\in D, u \in U.
 	\]
@@
 	\[
-	f(y,u) = \left(\varphi(y) + \sum_{j=1}^M{u_j}\right)\left(h_{min} + h(y,u)\right).
+	f(y,u) = \left(\varphi(y) + \sum_{j=1}^M{u_j}\right)\left(h_{max} + h(y,u)\right).
 	\]

### 2019-04-12 2019/RSD_Integer/RSD2019.tex
minor correction

@@
 
-\usepackage{hyperref}
+%\usepackage{hyperref}
 \usepackage{graphicx}
@@
 formulae
-\begin{eqnarray*}
-& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}, \; \nu(x_{t_j-1})\neq \nu(x_{t_j}), \\
-& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}- \frac{\mathrm{sign}(z_{t_j}-z_{t_j-
-1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \; \nu(x_{t_j-
-1})=\nu(x_{t_j})=\nu. \\
-\end{eqnarray*}
+\begin{gather*}
+x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}, \; \nu(x_{t_j-1})\neq \nu(x_{t_j}),\\
+x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}- \frac{\mathrm{sign}(z_{t_j}-z_{t_j-
+1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \\
+\nu(x_{t_j-1})=\nu(x_{t_j})=\nu.
+\end{gather*}
 
@@
 This generator of multiextremal functions is often used for the investigations of the global 
-optimization algorithms ~\cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}. 
+optimization algorithms ~\cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015,Barkalov2018}. 
 In Fig.~\ref{example} (a) and (b), the contour plots of two-dimensional GKLS functions are 
@@
 
+\bibitem{Barkalov2018}
+Barkalov, K., Strongin, R.: Solving a set of global optimization problems by the parallel technique with uniform convergence. J. Glob. Optim. \textbf{71}(1), 21--36 (2018)
+
 \end{thebibliography}

### 2019-04-12 2019/NUMTA_Barkalov/LNCS/numta19.tex
Version 0

@@
 \usepackage{graphicx}
-%\usepackage{hyperref}
+\usepackage{hyperref}
 % Used for displaying a sample figure. If possible, figure files should
@@
 
+\usepackage[utf8x]{inputenc}
+\usepackage[english,russian]{babel}
+
 \begin{document}
@@
 \title{Acceleration of Global Search through Dual Lipschitz Constant Estimates
-\thanks{This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.}}
+\thanks{This research was supported by the the Russian Science Foundation,
+project No.\,16-11-10150.}}
 %
@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-150--250 words.
+The paper considers global optimization problems with a black-box objective 
+function satisfying the Lipschitz condition. Efficient algorithms for this 
+class of problems require reliable estimates of the Lipschitz constant to be 
+introduced. The various approaches have been proposed to take into account both
+global and local properties of the objective function. In particular, algorithms
+using local estimates of the Lipschitz constant have shown their potential.
+The new approach presented in this paper is based on simultaneous use of two
+estimates: one is substantially larger then the other. 
+The larger estimate ensures global convergence and the smaller one reduces 
+the total number of trials needed to find the global optimizer.
+Results of numerical experiments on the random sample of multidimensional 
+functions demonstrate the efficiency of the suggested approach.
 
-\keywords{First keyword  \and Second keyword \and Another keyword.}
+\keywords{Global optimization \and Multiextremal problems 
+\and Lipschitz constant estimates}
 \end{abstract}
@@
 
-The paper considers global optimization problems with a black-box objective function satisfying the Lipschitz condition. 
-Efficient algorithms for this class of problems require reliable estimates of the Lipschitz constant to be introduced. 
-In fundamental work \cite{Strongin2000} various approaches have been proposed to take into account both global and local properties of the objective function. 
-%В частности, показали свою перспективность  алгоритмы с использованием локальных оценок константы Липшица 
-In particular, algorithms using local estimates of the Lipschitz constant have shown their potential \cite{Sergeyev2010,Sergeyev2016}.
+% Введение должно быть полностью новым
+% Константа Липшица - основное в методах глобальной оптимизации
+% Постановку задачи - дать во введении? Можно, ради экономии места
+% А в журнальной статье постановку задачи вынести в отельный раздел
 
-The new approach presented in this paper is based on simultaneous use of two estimates: one is substantially larger then the other. 
-The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
-Results of numerical experiments on the random sample of multidimensional functions demonstrate the efficiency of the suggested approach.
+\section{Global search algorithm}
+
+According to the global search algorithm, the first two trials are executed at 
+the points $y^0=y(0), y^1=y(1)$. The choice of the point $y^{k+1},k\geq 1,$  
+for the next $(k+1)$-th trial is defined by the following rules.
+
+\begin{enumerate}
+	\item 
+	Renumber the preimages of all the points $y^i=y(x^i)$
+	from the trials already performed  	
+%\begin{equation}\label{y_i} 
+%y^0=y(x^0), y^1=y(x^1),...,y^k=y(x^k)
+%\end{equation}
+by subscripts in the increasing order of their coordinates, i.e.
+\begin{equation}\label{x_i}
+0=x_0<x_1<\dots <x_k=1,
+\end{equation}
+and associate these with the values $z_i=\varphi(y(x_i)), 0\leq i \leq k,$ 
+computed at these points.
+\item
+Compute the maximum absolute value of the first divided differences
+\begin{equation}\label{mu}
+\mu = \max_{1 \leq i \leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
+\end{equation}
+where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$. If $\mu = 0$, set $\mu = 1$.
+\item
+For each interval $(x_{i-1}, x_i), \; 1\leq i \leq k,$  calculate the value
+\begin{equation}\label{R}
+R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r\mu}
+\end{equation}
+called the \textit{characteristic} of the interval; the real number $r>1$ being 
+the input parameter of the algorithm.
+\item 
+Select the interval $(x_{t-1},x_t)$ corresponding to the maximum characteristic
+\begin{equation}\label{MaxR}
+R(t)= \max_{1 \leq i \leq k}R(i).
+\end{equation}
+\item
+Carry out the next trial at the point $x^{k+1}\in(x_{t-1},x_t)$ calculated using
+the following formula
+\begin{equation}\label{xk1}
+x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}
+\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N.
+\end{equation}
+\end{enumerate}
+
+The algorithm terminates if the condition $\Delta_t < \epsilon$ is satisfied
+where $t$ is from (\ref{MaxR}), and $\epsilon>0$ is the predefined accuracy.
+
+The theory of convergence of this algorithm is provided in \cite{Strongin2000}.
+%The modifications taking into account .... are given in [....].
+
+\section{Алгоритм с двойной оценкой Lipschitz constant}
+
+Комментарий об использовании $\mu$ как оценки константы Липшица
+
+Ключевое условие из теорема о сходимости - $r\mu > 2^{3-1/N}L\sqrt{N+3} $
+
+Возьмем большое значение параметра $r$ - алгоритм гарантированно сойдется к глобальному минимум, но сделает много итераций, т.к. значение константы Липшица будет переоценено.
+
+Возьмем малое значение $r$ - алгоритм будет сходиться быстро, но может потерять свойство сходимости к глобальному экстремуму, т.к. значение константы Липшица будет недооценено.
+
+Характеристики $R(i)$ зависят от параметра $r$, и они будут несравнимы при разных значениях $r$.
+
+Сравнимость характеристик можно получить, если ввести нормировочные коэффициент (формула).
+
+\section{Numerical Experiments}
 

### 2019-04-17 2019/NUMTA_Barkalov/LNCS/numta19.tex

@@
 \usepackage{hyperref}
+\usepackage{amsmath}
+
 % Used for displaying a sample figure. If possible, figure files should
@@
 
-% Введение должно быть полностью новым
+The paper considers global optimization problems of the form 
+\begin{gather}
+ \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
+ D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\},
+\end{gather}
+where the objective function is a black-box function and assumed to satisfy the Lipschitz condition
+\[
+\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D,
+\]
+with the constant $L$ unknown a priori.
+
+Предположение липшицевости целевой функции является типичным для многих алгоритмов глобальной оптимизации (ссылки)
+
+%Здесь же - про редукцию размерности!
+% Или в конце введения? 
+% Или вообще в следующем разделе?
+
+При этом наиболее важной проблемой, так или иначе решаемой в данных алгоритмах, является адаптивная оценка неизвестной константы Липшица на основе полученной поисковой информации.
+
+Значение константы Липшица существенно влияет на скорость сходимости алгоритмов липшицевой глобальной оптимизации, поэтому столь важным является вопрос ее корректной оценки.
+Заниженная оценка инстинного значения константы Липшица может привести к потере сходимости алгоритма к глобальному решению.
+В то же время слишком большое значение оценки константы $L$ для целевой функции предполагает сложную структуру функции с резкими перепадами ее значений.
+Поэтому завышенная оценка константы L, не соответствующая истинному поведению функции, влечет за собой медленную сходимость алгоритма к точке глобального минимума.
+
+Известны несколько типичных способов адаптивного оценивания константы Липшица:
+1. оценивание глобальной константы Липшица во всей области поиска D (см., например, [] ). 
+2. оценивание локальных оценок констант Липшица $L_i$ в различных подобластях $D_i$ области поиска $D$ (см. ). 
+3. выбор оценок константы $L$ из некоторого множетсва возможных значений.
+
+%связка
+
+В данной работе предлагается алгоритм, который испольует две глобальных оценки константы Липшица, одна из которых значительно больше другой.
+
 % Константа Липшица - основное в методах глобальной оптимизации
-% Постановку задачи - дать во введении? Можно, ради экономии места
-% А в журнальной статье постановку задачи вынести в отельный раздел
 
@@
 
-Комментарий об использовании $\mu$ как оценки константы Липшица
+Алгоритм глобального поиска, изложенной в предыдущем разделе, предназначен для решения многоэкстремальных задач, в которых целевая функция удовлетворяет условию Липшица.
+При этом для для работы алгоритма не требуется задания  значений константы Липшица.
+Оценка константы осуществляется в процессе глобального поиска на основе имеющейся поисковой информации.
+В соответствии с теоремой \cite{Strongin2000}, последовательность точек испытаний $\{y^k\}$ будет сходиться к global minimizer $y^*$, если выполнено условие  
+\begin{equation}\label{cond}
+r\mu > 2^{3-1/N}L\sqrt{N+3}.
+\end{equation}
+Таким образом, подходящий выбор параметра $r$ из (\ref{R}) позволяет использовать значение $r\mu / 2^{3-1/N}\sqrt{N+3}$ как оценку константы Липшица для целевой функции $\varphi(y)$.
 
-Ключевое условие из теорема о сходимости - $r\mu > 2^{3-1/N}L\sqrt{N+3} $
+Выполнение условия (\ref{cond}) будет гарантировано при выборе большого значения параметра $r$, однако при этом метод выполнит большое количество испытаний до выполнения условия остановки.
+Выбор малого значения параметра $r$ (что соответствует заниженной оценке константы Липшица) значительно сократит число испытаний, но может нарушить сходимость к глобальному экстремуму.
 
-Возьмем большое значение параметра $r$ - алгоритм гарантированно сойдется к глобальному минимум, но сделает много итераций, т.к. значение константы Липшица будет переоценено.
-
-Возьмем малое значение $r$ - алгоритм будет сходиться быстро, но может потерять свойство сходимости к глобальному экстремуму, т.к. значение константы Липшица будет недооценено.
+Перспективным является подход, при котором в правилах алгоритма будут использованы две оценки константы Липшица, одна из которых будет значительно выше другой, что соответствует использованию в алгоритме двух парметров $r_{max}$ и $r_{min}$, где $r_{max} > r_{min}$.
 
@@
 Сравнимость характеристик можно получить, если ввести нормировочные коэффициент (формула).
+
+Таким образом, правила алгоритма с двойной оценкой константы Липшица будут полностью повторять правила алгоритма глобального поиска, за исключением правила 3 алгоритма.
+
+Новое правило вычисленя характеристики $R(i)$ интервала $(x_{i-1}, x_i)$ будет состоять из следующих действий:
+\begin{itemize}
+\item
+Вычислить значение $R_{max}(i)$, соответствующее большей оценке константы Липшица
+\[
+R_{max}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{max}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r_{max}\mu}.
+\]
+\item
+Вычислить значение $R_{min}(i)$, соответствующее меньшей оценке константы Липшица
+\[
+R_{min}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{min}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r_{min}\mu}.
+\]
+\item
+Определить характеристику интервала $R(i)$ как
+\[
+R(i) = \max\{\rho R_{min}(i),R_{max}(i)\}, где \rho = \left(\frac{1-1/r_{max}}{1-1/r_{min}}\right)^2.
+\]
+\end{itemize}
 

### 2019-04-17 2019/NUMTA_Barkalov/LNCS/numta19.tex

@@
 
-Предположение липшицевости целевой функции является типичным для многих алгоритмов глобальной оптимизации (ссылки)
+Предположение липшицевости целевой функции является типичным для многих подходов к разработке алгоритмов глобальной оптимизации (ссылки).
 
@@
 Заниженная оценка инстинного значения константы Липшица может привести к потере сходимости алгоритма к глобальному решению.
-В то же время слишком большое значение оценки константы $L$ для целевой функции предполагает сложную структуру функции с резкими перепадами ее значений.
-Поэтому завышенная оценка константы L, не соответствующая истинному поведению функции, влечет за собой медленную сходимость алгоритма к точке глобального минимума.
+В то же время слишком большое значение оценки константы $L$ для целевой функции, %предполагает сложную структуру функции с резкими перепадами ее значений.
+Поэтому завышенная оценка константы L,
+не соответствующее ее истинному поведению, влечет за собой медленную сходимость алгоритма к точке глобального минимума.
 
 Известны несколько типичных способов адаптивного оценивания константы Липшица:
-1. оценивание глобальной константы Липшица во всей области поиска D (см., например, [] ). 
-2. оценивание локальных оценок констант Липшица $L_i$ в различных подобластях $D_i$ области поиска $D$ (см. ). 
-3. выбор оценок константы $L$ из некоторого множетсва возможных значений.
+\begin{itemize}
+	\item оценивание глобальной константы Липшица во всей области поиска D (см., например, [] ).
+	\item оценивание локальных оценок констант Липшица $L_i$ в различных подобластях $D_i$ области поиска $D$ (см. ).
+	\item выбор оценок константы $L$ из некоторого множества возможных значений.
+\end{itemize}
 
 %связка
+Каждый из указанных подходов обладает своими достоинствами и недостатками. Например,
 
 В данной работе предлагается алгоритм, который испольует две глобальных оценки константы Липшица, одна из которых значительно больше другой. 
+The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
+
+% геометрическая интерпретация
+
+Выбор того, какая из двух оценок будет использоваться в правилах алгоритма, осуществляется адаптивно, в зависимости от поведения функции и фазы поиска.
+
+Строгое обоснование предложенного подхода выходит за рамки данной первоначальной публикации и будет сделано в последующих работах. 
+Здесь же приведены результаты вычислительных экспериментов, наглядно демонстрирующие эффективность предложенного алгоритма.
+При проведении вычислительных экспериментов было решено несколько сотен тестовых многоэкстремальных задач разной размерности.
 

### 2019-04-18 2019/NUMTA_Barkalov/LNCS/numta19.tex

@@
 \begin{gather}
- \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
- D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\},
+ \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problem}\\
+ D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\} \label{D},
 \end{gather}
@@
 % Или в конце введения? 
-% Или вообще в следующем разделе?
+% Или вообще в следующем разделе? - Да, в следующем разделе
 
@@
 
-% геометрическая интерпретация
+% геометрическая интерпретация ?
 
@@
 
-\section{Global search algorithm}
+\section{Global search algorithm and dimensionality reduction}
+
+Типичным способом конструирования алгоритмов глобальной оптимизации является адаптация эффективных алгоритмов для решения одномерных задач к решению многомерных задач, see, for example, the diagonal partitions method in [18] or
+the simplicial partitions method in [19].
+
+В данной статье используется подход, основанный на идеях редукции размерности с помощью Peano-Hilbert curves, which continuously and unambiguously maps the unit interval $[0,1]$ onto the $n$-dimensional cube $D$ from (\ref{D}).
+By using this kind of mapping it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem
+\[
+\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}},
+\]
+where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
+\[
+\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N},
+\]
+with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation
+$ H=2 L \sqrt{N+3}$.
+
+Let us call the process of computing a function value (including the construction of the image $y=y(x)$) a \textit{trial}, and the pair $(x, z = \varphi(y))$ the outcome of the trial.
+
+%связка - формулируем правила алгоритма глобального поиска
 
@@
 For each interval $(x_{i-1}, x_i), \; 1\leq i \leq k,$  calculate the value 
+$R(i)$ called the \textit{characteristic} of the interval
 \begin{equation}\label{R}
-R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r\mu}
+R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r\mu},
 \end{equation}
-called the \textit{characteristic} of the interval; the real number $r>1$ being 
+where 
+\begin{equation}\label{z}
+z^*= \min_{0\leq i\leq k}z_i
+\end{equation} 
+and the real number $r>1$ is 
 the input parameter of the algorithm.
@@
 
-Характеристики $R(i)$ зависят от параметра $r$, и они будут несравнимы при разных значениях $r$.
+Правила алгоритма с двойной оценкой константы Липшица будут полностью повторять правила алгоритма глобального поиска, за исключением правила 3 алгоритма.
 
-Сравнимость характеристик можно получить, если ввести нормировочные коэффициент (формула).
-
-Таким образом, правила алгоритма с двойной оценкой константы Липшица будут полностью повторять правила алгоритма глобального поиска, за исключением правила 3 алгоритма.
-
-Новое правило вычисленя характеристики $R(i)$ интервала $(x_{i-1}, x_i)$ будет состоять из следующих действий:
+Новое правило вычисления характеристики $R(i)$ интервала $(x_{i-1}, x_i)$ будет состоять из следующих действий:
 \begin{itemize}
@@
 Определить характеристику интервала $R(i)$ как
-\[
+\begin{equation}\label{pho}
 R(i) = \max\{\rho R_{min}(i),R_{max}(i)\}, где \rho = \left(\frac{1-1/r_{max}}{1-1/r_{min}}\right)^2.
-\]
+\end{equation}
 \end{itemize}
 
+Указанный способ вычисления характеристики имеет следующее обоснование. 
+Пусть текущее минимальное значение $z^*$ из (\ref{z}) достигается на в левой точке $i$-го интервала, т.е. $z^* = z_{i-1}$. Тогда, в соответствии с правилом (\ref{R}), будут верно неравенство
+\[
+R(i) \geq \Delta_i \left( 1 - \frac{1}{r} \right)^2.
+\]
+Следовательно, две нижние оценки для двух характеристик текущего лучшего интервала $R_{min}(i)$ и $R_{max}(i)$, вычисленных с разными параметрами $r_{min}$ и $r_{max}$, будут иметь разные значения.
+Однако если первую из них домножить на коэффициент $\rho$ из (\ref{pho}), то нижние оценки характеристик лучшего интервала будут совпадать, что сделает сравнимыми и характеристики, вычисленные для остальных интервалов.
+
 \section{Numerical Experiments}
+
+The numerical comparison of the algorithms has been carried out using the GKLS 
+test problem generator \cite{Gaviano2003}. 
+This generator of multiextremal functions is often used for the investigations of the global 
+optimization algorithms ~\cite{Paulavicius2014,Sergeyev2015,%Lebedev2015,Gergel2015,
+Barkalov2018}.
+%This generator allows multiextremal optimization problems to be generated with known properties (the number of localminima, the size of their domains of attraction, the global minimizer, etc.). 
+Six GKLS classes of differentiable test functions of the dimensions $N = 3,4,5$
+have been used. For each dimension, both \textit{Hard} and \textit{Simple}
+classes have been considered. The difficulty of a class was increased either by
+decreasing the radius of the attraction region of the global minimizer, or by
+decreasing the distance from the global minimizer $y^\ast$ to the domain
+boundaries. The global minimizer $y^\ast$ was considered to be found, if the
+algorithm generated a trial point $y^k$ in the vicinity of the global minimum,
+i.e. $\left|y^k-y^\ast\right| <\delta\left\|b-a\right\|$, where $\delta = 0.01$,
+$a$ and $b$ are the boundaries of the search domain $D$.
+
+%УКАЗАТЬ ПРАВИЛЬНЫЕ ПАРАМЕТРЫ
+When using the Global Search Algorithm, the parameter $r$ was set to $5$, 
+the evolvent construction parameter was fixed as $m = 10$. 
+The maximum allowable number of iterations per problem was $K_{max} = 10^6$.
+
+Каждая серия задач была решена с помощью исходного алгоритма глобального поиска (GSA) и спомощью метода с двойной оценкой константы Липшица (GSA-DL).
+Для GSA использовалось минимальное (с шагом 0.1) значение параметра $r = 5.6$, при котором успешно решились все задачи.
+Для GSA-DL использовались значения $r_{max} = 5.6$ и $r_{min} = 2.0$.
+
+%Структуру таблицы можно взять из статьи со Стронгиным
+
+%\begin{table}
+	%\caption{Average number of iterations for the \textit{Simple} class}
+	%\label{tab:1}
+	%\center
+	%\begin{tabular}{cccccccccccc}
+		%\hline\noalign{\smallskip}
+		%$p$ & \multicolumn{2}{c}{ $N=2$ } & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
+		%\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12}  \noalign{\smallskip}
+		 %& $S1$ & $S2$ & & $S1$ & $S2$ & & $S1$ & $S2$ & & $S1$ & $S2$  \\
+		%\noalign{\smallskip} \hline \noalign{\smallskip}
+%1	&	408	&	408	&	&	2502	&	2502	&	&	28254	&	28254	&	&	87261	&	87261	\\
+%2	&	408	&	415	&	&	2502	&	2496	&	&	28254	&	27206	&	&	87261	&	85994	\\
+%4	&	408	&	405	&	&	2502	&	2430	&	&	28254	&	30121	&	&	87261	&	78064	\\
+%8	&	408	&	402	&	&	2502	&	2505	&	&	28254	&	28980	&	&	87261	&	84337	\\
+%16	&	408	&	415	&	&	2502	&	2501	&	&	28254	&	28206	&	&	87261	&	82564	\\
+%24	&	408	&	400	&	&	2502	&	2416	&	&	28254	&	27868	&	&	87261	&	86252	\\
+		%\noalign{\smallskip}\hline
+	%\end{tabular}
+%\end{table}
+
+Превосходство алгоритма GSA-DL над его прототипом наглядно подтверждается также и операционными характеристиками алгоритмов.
+
+Результат решения всех задач серии представляется функцией $p(k)$, характеризующей долю задач, решенных за $k$ итераций. Такую функцию будем называть операционной характеристикой алгоритма.
+
+Операционные характеристики для обоих методов, полученные на сериях пятимерных задач GKLS Simple и GKLS Hard представлены соответственно на рис.3 и 4.
+При этом нижние (непрерывные) кривые на обеих фигурах характеризуют GSA, а верхние (разрывные) линии — GSA-DL.
+Указанное расположение кривых показывает, что алгоритм с двумя оценками константы Липшица обеспечивает в среднем более быстрое решение задач данного класса, чем алгоритм с использованием одной глобальной оценки константы Липшица.
 

### 2019-04-25 2019/NUMTA_Barkalov/LNCS/numta19.tex

@@
 The new approach presented in this paper is based on simultaneous use of two
-estimates: one is substantially larger then the other. 
+estimates: one is substantially larger than the other. 
 The larger estimate ensures global convergence and the smaller one reduces 
@@
 \begin{itemize}
-	\item оценивание глобальной константы Липшица во всей области поиска D (см., например, [] ).
-	\item оценивание локальных оценок констант Липшица $L_i$ в различных подобластях $D_i$ области поиска $D$ (см. ).
-	\item выбор оценок константы $L$ из некоторого множества возможных значений.
+	\item оценивание глобальной константы Липшица во всей области поиска $D$ \cite{Horst1996,Pinter1996,Strongin2000}.
+	\item оценивание локальных оценок констант Липшица $L_i$ в различных подобластях $D_i$ области поиска $D$ \cite{Kvasov2003,Sergeyev2010,Sergeyev2016}.
+	\item выбор оценок константы $L$ из некоторого множества возможных значений \cite{Jones1993,Gablonsky2001,Jones2009,Sergeyev2006}.
 \end{itemize}
 
-%связка
-Каждый из указанных подходов обладает своими достоинствами и недостатками. Например,
+%Каждый из указанных подходов обладает своими достоинствами и недостатками.
+Например, использование только глобальной константы Липшица во всей области поиска может замедлить сходимость алгоритма к точке глобального минимума.
+А использование локальных оценок, ускоряющих сходимость, требует адекватной настройки параметров алгоритма для сохранения глобальной сходимости.
 
-В данной работе предлагается алгоритм, который испольует две глобальных оценки константы Липшица, одна из которых значительно больше другой. 
+В данной работе рассмотрен алгоритм, котором предлагается использовать две глобальных оценки константы Липшица, одна из которых значительно больше другой. 
 The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
-
 % геометрическая интерпретация ?
-
 Выбор того, какая из двух оценок будет использоваться в правилах алгоритма, осуществляется адаптивно, в зависимости от поведения функции и фазы поиска.
@@
 Строгое обоснование предложенного подхода выходит за рамки данной первоначальной публикации и будет сделано в последующих работах.
-Здесь же приведены результаты вычислительных экспериментов, наглядно демонстрирующие эффективность предложенного алгоритма.
+Здесь же приведем результаты вычислительных экспериментов, наглядно демонстрирующие эффективность предложенного алгоритма.
 При проведении вычислительных экспериментов было решено несколько сотен тестовых многоэкстремальных задач разной размерности.
-
-% Константа Липшица - основное в методах глобальной оптимизации
 
@@
 
-\section{Алгоритм с двойной оценкой Lipschitz constant}
+\section{Algorithm with Dual Lipschitz Constant Estimates}
 
 Алгоритм глобального поиска, изложенной в предыдущем разделе, предназначен для решения многоэкстремальных задач, в которых целевая функция удовлетворяет условию Липшица.
-При этом для для работы алгоритма не требуется задания  значений константы Липшица.
+При этом для для работы алгоритма не требуется задания  значений константы.
 Оценка константы осуществляется в процессе глобального поиска на основе имеющейся поисковой информации.
@@
 
-Перспективным является подход, при котором в правилах алгоритма будут использованы две оценки константы Липшица, одна из которых будет значительно выше другой, что соответствует использованию в алгоритме двух парметров $r_{max}$ и $r_{min}$, где $r_{max} > r_{min}$.
+Перспективным является подход, при котором в правилах алгоритма будут использованы две оценки константы Липшица, одна из которых будет значительно выше другой, что соответствует использованию в алгоритме двух парметров $r_{glob}$ и $r_{loc}$, где $r_{glob} > r_{loc}$.
 
@@
 \item
-Вычислить значение $R_{max}(i)$, соответствующее большей оценке константы Липшица
+Вычислить значение $R_{glob}(i)$, соответствующее большей оценке константы Липшица
 \[
-R_{max}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{max}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r_{max}\mu}.
+R_{glob}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{glob}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r_{glob}\mu}.
 \]
 \item
-Вычислить значение $R_{min}(i)$, соответствующее меньшей оценке константы Липшица
+Вычислить значение $R_{loc}(i)$, соответствующее меньшей оценке константы Липшица
 \[
-R_{min}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{min}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r_{min}\mu}.
+R_{loc}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{loc}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r_{loc}\mu}.
 \]
@@
 \begin{equation}\label{pho}
-R(i) = \max\{\rho R_{min}(i),R_{max}(i)\}, где \rho = \left(\frac{1-1/r_{max}}{1-1/r_{min}}\right)^2.
+R(i) = \max\{\rho R_{loc}(i),R_{glob}(i)\}, где \rho = \left(\frac{1-1/r_{glob}}{1-1/r_{loc}}\right)^2.
 \end{equation}
@@
 \]
-Следовательно, две нижние оценки для двух характеристик текущего лучшего интервала $R_{min}(i)$ и $R_{max}(i)$, вычисленных с разными параметрами $r_{min}$ и $r_{max}$, будут иметь разные значения.
+Следовательно, две нижние оценки для двух характеристик текущего лучшего интервала $R_{loc}(i)$ и $R_{glob}(i)$, вычисленных с разными параметрами $r_{loc}$ и $r_{glob}$, будут иметь разные значения.
 Однако если первую из них домножить на коэффициент $\rho$ из (\ref{pho}), то нижние оценки характеристик лучшего интервала будут совпадать, что сделает сравнимыми и характеристики, вычисленные для остальных интервалов.
@@
 Для GSA использовалось минимальное (с шагом 0.1) значение параметра $r = 5.6$, при котором успешно решились все задачи.
-Для GSA-DL использовались значения $r_{max} = 5.6$ и $r_{min} = 2.0$.
+Для GSA-DL использовались значения $r_{glob} = 5.6$ и $r_{loc} = 2.0$.
 
-%Структуру таблицы можно взять из статьи со Стронгиным
+%Структуру таблицы можно взять из статьи
 
@@
 
-Результат решения всех задач серии представляется функцией $p(k)$, характеризующей долю задач, решенных за $k$ итераций. Такую функцию будем называть операционной характеристикой алгоритма.
+Результат решения всех задач серии будем представлять функцией $p(k)$, характеризующей долю задач, решенных за $k$ итераций.
+Такую функцию будем называть \textit{operational characteristic} of the algorithm..
 
-Операционные характеристики для обоих методов, полученные на сериях пятимерных задач GKLS Simple и GKLS Hard представлены соответственно на рис.3 и 4.
-При этом нижние (непрерывные) кривые на обеих фигурах характеризуют GSA, а верхние (разрывные) линии — GSA-DL.
-Указанное расположение кривых показывает, что алгоритм с двумя оценками константы Липшица обеспечивает в среднем более быстрое решение задач данного класса, чем алгоритм с использованием одной глобальной оценки константы Липшица.
+Операционные характеристики для методов GSA и GSA-DL, полученные при решении серий задач GKLS Simple и GKLS Hard размерности $N=4$ представлены соответственно на рис. \ref{oper}.
+
+\begin{figure}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklss_4.png} \\ (a)}
+\end{minipage}
+\hfill
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklsh_4.png} \\ (b)}
+\end{minipage}
+\caption{Operational characteristics on GKLS Simple (a) and Hard (b) classes.}
+\label{oper}
+\end{figure}
+
+Нижние кривые на обеих фигурах характеризуют GSA, а верхние линии — GSA-DL.
+Указанное расположение кривых показывает, что алгоритм с двумя оценками константы Липшица обеспечивает в среднем значительно более быстрое решение задач данного класса, чем алгоритм с использованием единственной оценки константы Липшица.
+
+При этом при решении задач класса Simple в GSA использовался параметр $r=4.8$, в GSA-DL использовались параметры $r_{glob}=4.8$ and $r_{loc}=1.8$; для класса Hard в GSA использовался параметр $r=5.6$, в GSA-DL использовались параметры $r_{glob}=5.6$ and $r_{loc}=2.4$.
 

### 2019-04-26 2019/NUMTA_Barkalov/LNCS/numta19.tex

@@
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Acceleration of Global Search}
 % If the paper title is too long for the running head, you can set
@@
 
-Предположение липшицевости целевой функции является типичным для многих подходов к разработке алгоритмов глобальной оптимизации (ссылки).
-
-%Здесь же - про редукцию размерности!
-% Или в конце введения? 
-% Или вообще в следующем разделе? - Да, в следующем разделе
-
+Предположение липшицевости целевой функции является типичным для многих подходов к разработке алгоритмов глобальной оптимизации \cite{Pinter1996,Strongin2000,Zilinskas2010,Evtushenko2013}.
 При этом наиболее важной проблемой, так или иначе решаемой в данных алгоритмах, является адаптивная оценка неизвестной константы Липшица на основе полученной поисковой информации.
@@
 \begin{itemize}
-	\item оценивание глобальной константы Липшица во всей области поиска $D$ \cite{Horst1996,Pinter1996,Strongin2000}.
-	\item оценивание локальных оценок констант Липшица $L_i$ в различных подобластях $D_i$ области поиска $D$ \cite{Kvasov2003,Sergeyev2010,Sergeyev2016}.
+	\item глобальная оценка константы Липшица $L$ во всей области поиска $D$ \cite{Horst1996,Pinter1996,Strongin2000}.
+	\item локальные оценки константы Липшица $L_i$ в различных подобластях $D_i$ области поиска $D$ \cite{Kvasov2003,Sergeyev2010,Sergeyev2016}.
 	\item выбор оценок константы $L$ из некоторого множества возможных значений \cite{Jones1993,Gablonsky2001,Jones2009,Sergeyev2006}.
@@
 
-%Каждый из указанных подходов обладает своими достоинствами и недостатками.
-Например, использование только глобальной константы Липшица во всей области поиска может замедлить сходимость алгоритма к точке глобального минимума.
+Каждый из указанных подходов обладает своими достоинствами и недостатками.
+Например, использование только глобальной оценки во всей области поиска может замедлить сходимость алгоритма к точке глобального минимума.
 А использование локальных оценок, ускоряющих сходимость, требует адекватной настройки параметров алгоритма для сохранения глобальной сходимости.
@@
 The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
-% геометрическая интерпретация ?
 Выбор того, какая из двух оценок будет использоваться в правилах алгоритма, осуществляется адаптивно, в зависимости от поведения функции и фазы поиска.
@@
 Строгое обоснование предложенного подхода выходит за рамки данной первоначальной публикации и будет сделано в последующих работах.
-Здесь же приведем результаты вычислительных экспериментов, наглядно демонстрирующие эффективность предложенного алгоритма.
+Здесь же приведем результаты вычислительных экспериментов, наглядно демонстрирующие эффективность нового алгоритма.
 При проведении вычислительных экспериментов было решено несколько сотен тестовых многоэкстремальных задач разной размерности.
@@
 
-Типичным способом конструирования алгоритмов глобальной оптимизации является адаптация эффективных алгоритмов для решения одномерных задач к решению многомерных задач, see, for example, the diagonal partitions method in [18] or
-the simplicial partitions method in [19].
+Типичным способом конструирования алгоритмов глобальной оптимизации является адаптация эффективных алгоритмов для решения одномерных задач к решению многомерных задач, see, for example, the diagonal partitions method in \cite{Sergeyev2006} or the simplicial partitions method in \cite{Zilinskas2008}.
 
-В данной статье используется подход, основанный на идеях редукции размерности с помощью Peano-Hilbert curves, which continuously and unambiguously maps the unit interval $[0,1]$ onto the $n$-dimensional cube $D$ from (\ref{D}).
+В данной статье используется подход, основанный на идеях редукции размерности с помощью Peano-Hilbert curves \cite{Strongin2000,Sergeyev2013}, which continuously and unambiguously maps the unit interval $[0,1]$ onto the $n$-dimensional cube $D$ from (\ref{D}).
 By using this kind of mapping it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem
@@
 
-Let us call the process of computing a function value (including the construction of the image $y=y(x)$) a \textit{trial}, and the pair $(x, z = \varphi(y))$ the outcome of the trial.
+Let us call the process of computing a function value (including the construction of the image $y=y(x)$) a \textit{trial}, and the pair $\{x, z = \varphi(y(x))\}$ the outcome of the trial.
 
-%связка - формулируем правила алгоритма глобального поиска
-
-According to the global search algorithm, the first two trials are executed at 
+The global search algorithm (according to \cite{Strongin2000}) can be formulated as follows.
+The first two trials are executed at 
 the points $y^0=y(0), y^1=y(1)$. The choice of the point $y^{k+1},k\geq 1,$  
@@
 
-Правила алгоритма с двойной оценкой константы Липшица будут полностью повторять правила алгоритма глобального поиска, за исключением правила 3 алгоритма.
+Правила алгоритма с двойной оценкой константы Липшица будут полностью повторять алгоритм глобального поиска, за исключением правила 3 (вычисление характеристики).
 
@@
 \begin{equation}\label{pho}
-R(i) = \max\{\rho R_{loc}(i),R_{glob}(i)\}, где \rho = \left(\frac{1-1/r_{glob}}{1-1/r_{loc}}\right)^2.
+R(i) = \max\{\rho R_{loc}(i),R_{glob}(i)\}, где \rho = \left(\frac{1-1/r_{glob}}{1-1/r_{loc}}\right)^2,
 \end{equation}
-\end{itemize}
+и зафиксировать значение $r = r_{loc}$, если  $\rho R_{loc}(i) > R_{glob}(i)$, иначе зафиксировать $r=r_{glob}$.
+Данное значение $r$ будет использовано в правиле 5 алгоритма при вычислении точки нового испытания.   
+\end{itemize}.
 
 Указанный способ вычисления характеристики имеет следующее обоснование. 
-Пусть текущее минимальное значение $z^*$ из (\ref{z}) достигается на в левой точке $i$-го интервала, т.е. $z^* = z_{i-1}$. Тогда, в соответствии с правилом (\ref{R}), будут верно неравенство
+Пусть текущее минимальное значение $z^*$ из (\ref{z}) достигается в левой точке $i$-го интервала, т.е. $z^* = z_{i-1}$. Тогда, в соответствии с правилом (\ref{R}), будут верно неравенство
 \[
@@
 Каждая серия задач была решена с помощью исходного алгоритма глобального поиска (GSA) и спомощью метода с двойной оценкой константы Липшица (GSA-DL).
-Для GSA использовалось минимальное (с шагом 0.1) значение параметра $r = 5.6$, при котором успешно решились все задачи.
-Для GSA-DL использовались значения $r_{glob} = 5.6$ и $r_{loc} = 2.0$.
-
+Среднее число итераций, выполненное алгоритмами, отражено в табл.
+\ref{tab:1}.
+При этом при решении задач класса Simple в GSA использовалось единственное значение параметра $r=4.8$, при решении задач класса Hard -- $r=5.6$.
+Указанное значение является минимальным (с точностью 0.1) значением, при котором успешно решились все задачи.
+При решении данных классов задач с использованием GSA-DL указанное ранее значение параметра $r$ выбиралось для верхней оценки константы Липшица, т.е.
+задавалось значение $r_{glob} = r$, которое дополнялось значениями $r_{loc}=1.8$ и $r_{loc}=2.4$.
 %Структуру таблицы можно взять из статьи
 
-%\begin{table}
-	%\caption{Average number of iterations for the \textit{Simple} class}
-	%\label{tab:1}
-	%\center
-	%\begin{tabular}{cccccccccccc}
-		%\hline\noalign{\smallskip}
-		%$p$ & \multicolumn{2}{c}{ $N=2$ } & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
-		%\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12}  \noalign{\smallskip}
-		 %& $S1$ & $S2$ & & $S1$ & $S2$ & & $S1$ & $S2$ & & $S1$ & $S2$  \\
-		%\noalign{\smallskip} \hline \noalign{\smallskip}
-%1	&	408	&	408	&	&	2502	&	2502	&	&	28254	&	28254	&	&	87261	&	87261	\\
-%2	&	408	&	415	&	&	2502	&	2496	&	&	28254	&	27206	&	&	87261	&	85994	\\
-%4	&	408	&	405	&	&	2502	&	2430	&	&	28254	&	30121	&	&	87261	&	78064	\\
-%8	&	408	&	402	&	&	2502	&	2505	&	&	28254	&	28980	&	&	87261	&	84337	\\
-%16	&	408	&	415	&	&	2502	&	2501	&	&	28254	&	28206	&	&	87261	&	82564	\\
-%24	&	408	&	400	&	&	2502	&	2416	&	&	28254	&	27868	&	&	87261	&	86252	\\
-		%\noalign{\smallskip}\hline
-	%\end{tabular}
-%\end{table}
+\begin{table}
+	\caption{Average number of iterations}
+	\label{tab:1}
+	\center
+	\begin{tabular}{ccccccccc}
+		\hline\noalign{\smallskip}
+		$r$ & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
+		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \noalign{\smallskip}
+		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
+		\noalign{\smallskip} \hline \noalign{\smallskip}
+									AGS	&	2502	&	2502	&	&	28254	&	28254	&	&	87261	&	87261	\\
+AGS-DL, $r_{loc}=2.4$	&	2502	&	2496	&	&	28254	&	27206	&	&	87261	&	85994	\\
+AGS-DL, $r_{loc}=1.8$	&	2502	&	2430	&	&	28254	&	30121	&	&	87261	&	78064	\\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
 
@@
 
-Операционные характеристики для методов GSA и GSA-DL, полученные при решении серий задач GKLS Simple и GKLS Hard размерности $N=4$ представлены соответственно на рис. \ref{oper}.
+Операционные характеристики для методов GSA и GSA-DL, полученные при решении серий задач \textit{Simple} and \textit{Hard} размерностей $N=4$ и $N=5$ представлены соответственно на рис.
+\ref{oper4} и \ref{oper5}.
 
@@
 \end{minipage}
-\caption{Operational characteristics on GKLS Simple (a) and Hard (b) classes.}
-\label{oper}
+\caption{Operational characteristics for GKLS Simple (a) and Hard (b) classes, $N=4$.}
+\label{oper4}
+\end{figure}
+
+\begin{figure}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklss_4.png} \\ (a)}
+\end{minipage}
+\hfill
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklsh_4.png} \\ (b)}
+\end{minipage}
+\caption{Operational characteristics for GKLS Simple (a) and Hard (b) classes, $N=5$.}
+\label{oper5}
 \end{figure}
@@
 Указанное расположение кривых показывает, что алгоритм с двумя оценками константы Липшица обеспечивает в среднем значительно более быстрое решение задач данного класса, чем алгоритм с использованием единственной оценки константы Липшица.
-
-При этом при решении задач класса Simple в GSA использовался параметр $r=4.8$, в GSA-DL использовались параметры $r_{glob}=4.8$ and $r_{loc}=1.8$; для класса Hard в GSA использовался параметр $r=5.6$, в GSA-DL использовались параметры $r_{glob}=5.6$ and $r_{loc}=2.4$.
 

### 2019-04-30 2019/PAVT/pavt_paper.tex
PAVT: final submission
PAVT: update the final submission

@@
-%%%%%%%%%%%%%%%%%%%% author.tex
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-%%%%%%%%%%%%%%%% Springer
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %
-% RECOMMENDED
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
+% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 \usepackage{marvosym}
-\usepackage{amssymb}
-\usepackage{cite}
-\usepackage{subfig}
+
 % to typeset URLs, URIs, and DOIs
@@
 \usepackage{hyperref}
-\usepackage{caption}
-\captionsetup[table]{skip=9pt}
 \def\UrlFont{\rmfamily}
@@
 
+\usepackage{amssymb}
+\usepackage{cite}
+\usepackage{subfig}
+\usepackage{caption}
+\captionsetup[table]{skip=9pt}
+\captionsetup{labelfont=bf}
+
 \begin{document}
@@
 %
-\title{ComparisonOfSeveralSequentialAndParallelDerivative-freeGlobalOptimizationAlgorithms}
+\title{Comparison Of Several Sequential And Parallel Derivative-free Global Optimization Algorithms}
 %
@@
 Globalizer solver, which is developed at the Lobachevsky State University. The Globalizer is
-designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows
+designed to solve problems with black-box objective functions satisfying the Lipschitz condition and shows
 competitive performance with other similar solvers. The comparison is done on several sets of
@@
 thousand-dimensional space. In local optimization applying of the simplest gradient descent
-method may appear to be enough, whereas in order to \textit{guarantee} finding the global
+method may appear to be enough, whereas in order to guarantee finding the global
 optimum, an optimization method has to accumulate information about behavior of the
@@
 \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}. Nowadays, various stochastic global
-optimization algorithms have attracted much attention, first of all, the evolution ones
+optimization algorithms have attracted much attention, first of all, the evolutionary ones
 \cite{Storn1997, SCHLUTER2009, KennedyEberhart1995}. They have rather simple structure
@@
 
-The stopping condition, which terminated the trials, is defined by the inequality
+The stopping condition, which terminates the trials, is defined by the inequality
 $\rho_{t_j}<\varepsilon,\: j=\overline{1,p}$
@@
 
-Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the
+Using the multiple mappings allows solving initial problem (\ref{eq:task}) by parallel solving the
 problems
@@
 in the form of the tuples: the processor number \(s\), the trial point \(x{}^{k_s}\).
-The modifications of the method taking into account multiply criteria are given in \cite{GERGEL2017,Gergel2018}.
+The modifications of the method taking into account multiple criteria are given in \cite{GERGEL2017,Gergel2018}.
 \par
@@
 MIDACO (Mixed Integer Distributed Ant Colony Optimization) \cite{Schlueter2012} is a solver
-which implements the evolutionary algorithm
-the ant colony optimization metaheuristic for continuous search domains. ACO was originally
-proposed to solve the TSP, but later has been successfully adopted to solve mixed integer nonlinear
+which implements the evolutionary algorithm the ant colony optimization metaheuristic
+for continuous search domains. ACO was originally
+proposed to solve the Traveling Salesman Problem, but later has been successfully adopted to solve mixed integer nonlinear
 programming problems \cite{SCHLUTER2009}.
@@
 DIRECT which pays less attention to non-promising boxes and therefore
-  has less exploration power: it can converge faster on problems with few local minima, but lost the
-global one in complicated cases.
+  has less exploration power: it can converge faster on problems with few local minima, but may lose the
+global minimum in complicated cases.
 
@@
 algorithm to
-  the continuous search domain.
+  the continuous search domains.
 
@@
 CRS in not an
-  evolutional algorithm, although stores something like population and performs transformation
+  evolutionary algorithm, although stores something like population and performs transformation
 resembling a mutation.
@@
 Let us denote the problem set obtained with the use of the first generator from \cite{grishaginClass}
-as \(F_{GR}\). The mechanism of generation of the problems \(F_{GR}\) doesn't provide the
+as \(F_{GR}\). The mechanism of generation of the problems \(GR\) doesn't provide the
 control of the problem complexity and of the number of local optima. However, the generated
 functions are known to be the multiextremal ones essentially. Besides, the problems generated by
-\(F_{GR}\) are the two-dimensional ones. In the present work, we will use 100 functions from the
+\(GR\) are the two-dimensional ones. In the present work, we will use 100 functions from the
 class \(F_{GR}\) generated randomly.
@@
 
-If one considers the costs of parallelism to be negligible as compared to the costs of computing the
+If one considers the costs of parallelization to be negligible as compared to the costs of computing the
 objective functions in the optimization problems, the speedup in time due to the use of the parallel
 method would be equal to the speedup with respect to the number of iterations. However, actually
-this assumption is true if the computing of the objective function is well computation costly only. In
+this assumption is true if the computing of the objective function is highly costly. In
 all numerical experiments, the time of computing the objective function was approximately $2.5\times
@@
 MIDACO decreased with increasing number of nodes on both problem classes whereas the
-reliability of Globalizer remains stable.
+reliability of Globalizer remains almost the same.
 
@@
 
+%
 % ---- Bibliography ----

### 2019-04-30 2019/PAVT/pavt_paper.tex
PAVT: update table with speedups
PAVT: final fixes

@@
 MIDACO has solved a greater number of problems in the parallel mode than in the sequential one,
-hence, has worked until the limit of trials is expired on a less number of problems.
+hence, it has performed much less iterations than in sequential mode.
 
@@
     & \textit{1}   & 25270 (27.3s) & 55180 (61.8s) & & 27645 (32.2s) & 72068 (132.5s)  \\
-  &  & \textit{16} & 14.3(11.7) & 14.9(12.6)  & &  16.9(14.4) & 16.7(14.4) \\
+  &  & \textit{16} & 14.3x(11.7x) & 14.9x(12.6x)  & &  16.9x(14.4x) & 16.7x(14.4x) \\
   \hline \noalign{\smallskip}
 II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
-  & \textit{1}      &   1.9(1.9) & 2.4(2.2)  & & 2.6(1.7) & 2.7(2.3) \\
-  &   & \textit{16} & 34.5(20.4) & 31.4(18.8) & & 24.5(18.8) & 32.0(29.1) \\
+  & \textit{1}      &   1.9x(1.9x) & 2.4x(2.2x)  & & 2.6x(1.7x) & 2.7x(2.3x) \\
+  &   & \textit{16} & 34.5x(20.4x) & 31.4x(18.8x) & & 24.5x(18.8x) & 32.0x(29.1x) \\
     \noalign{\smallskip}\hline	\noalign{\smallskip}
 III  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
-& \textit{1}      & 5.0(4.6) & 4.3(4.1) & & 4.8(3.9) & 4.5(4.4) \\
-&   & \textit{16} & 68.8(23.7) & 71.2(25.8) & & 52.3(32.9) & 57.0(50.2) \\
+& \textit{1}      & 5.0x(4.6x) & 4.3x(4.1x) & & 4.8x(3.9x) & 4.5x(4.4x) \\
+&   & \textit{16} & 68.8x(23.7x) & 71.2x(25.8x) & & 52.3x(32.9x) & 57.0x(50.2x) \\
   \noalign{\smallskip}\hline	\noalign{\smallskip}
 VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-& \textit{1}    & 12.0(8.7) & 10.9(8.3) & & 9.7(8.2)    & 7.3(9.0)    \\
-& & \textit{16} & 174.1(4.7) & 177.8(5.3) & & 101.6(51.6) & 93.1(84.1) \\
+& \textit{1}    & 12.0x(8.7x) & 10.9x(8.3x) & & 9.7x(8.2x)    & 7.3x(9.0x)    \\
+& & \textit{16} & 174.1x(4.7x) & 177.8x(5.3x) & & 101.6x(51.6x) & 93.1x(84.1x) \\
     \noalign{\smallskip}\hline
     V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
-    & \textit{1}    & 21.9(11.4)  & 23.0(12.8)  & & 12.4(11.1)  & 12.0(14.8)  \\
-    & & \textit{16} & 333.5(2.7) & 347.9(3.0) & & 164.5(82.0) & 183.2(129.7) \\
+    & \textit{1}    & 21.9x(11.4x)  & 23.0x(12.8x)  & & 12.4x(11.1x)  & 12.0x(14.8x)  \\
+    & & \textit{16} & 333.5x(2.7x) & 347.9x(3.0x) & & 164.5x(82.0x) & 183.2x(129.7x) \\
         \noalign{\smallskip}\hline

### 2019-05-10 2019/NUMTA_Barkalov/LNCS/numta19.tex
Before translation

@@
 %
-\title{Acceleration of Global Search through Dual Lipschitz Constant Estimates
-\thanks{This research was supported by the the Russian Science Foundation,
-project No.\,16-11-10150.}}
+\title{Acceleration of Global Search through Dual Lipschitz Constant Estimates\thanks{This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.}}
 %
@@
 
-Предположение липшицевости целевой функции является типичным для многих подходов к разработке алгоритмов глобальной оптимизации \cite{Pinter1996,Strongin2000,Zilinskas2010,Evtushenko2013}.
-При этом наиболее важной проблемой, так или иначе решаемой в данных алгоритмах, является адаптивная оценка неизвестной константы Липшица на основе полученной поисковой информации.
+%Russian
+Предположение липшицевости целевой функции является типичным для многих подходов к разработке алгоритмов глобальной оптимизации \cite{Evtushenko2013,Zilinskas2010,Pinter1996,Strongin2000}.
+При этом наиболее важной проблемой, так или иначе решаемой в данных алгоритмах, является адаптивная оценка неизвестной Lipschitz constant на основе полученной поисковой информации.
 
-Значение константы Липшица существенно влияет на скорость сходимости алгоритмов липшицевой глобальной оптимизации, поэтому столь важным является вопрос ее корректной оценки.
-Заниженная оценка инстинного значения константы Липшица может привести к потере сходимости алгоритма к глобальному решению.
-В то же время слишком большое значение оценки константы $L$ для целевой функции, %предполагает сложную структуру функции с резкими перепадами ее значений.
-Поэтому завышенная оценка константы L,
-не соответствующее ее истинному поведению, влечет за собой медленную сходимость алгоритма к точке глобального минимума.
+Значение Lipschitz constant существенно влияет на скорость сходимости алгоритмов липшицевой глобальной оптимизации, поэтому столь важным является вопрос ее корректной оценки.
+Заниженная оценка инстинного значения константы может привести к потере сходимости алгоритма к глобальному решению.
+В то же время слишком большое значение оценки константы $L$ для целевой функции, не соответствующее ее истинному поведению, влечет за собой медленную сходимость алгоритма к точке глобального минимума.
 
-Известны несколько типичных способов адаптивного оценивания константы Липшица:
+Известны несколько типичных способов адаптивного оценивания Lipschitz constant:
 \begin{itemize}
-	\item глобальная оценка константы Липшица $L$ во всей области поиска $D$ \cite{Horst1996,Pinter1996,Strongin2000}.
-	\item локальные оценки константы Липшица $L_i$ в различных подобластях $D_i$ области поиска $D$ \cite{Kvasov2003,Sergeyev2010,Sergeyev2016}.
-	\item выбор оценок константы $L$ из некоторого множества возможных значений \cite{Jones1993,Gablonsky2001,Jones2009,Sergeyev2006}.
+	\item глобальная оценка константы $L$ во всей области поиска $D$ \cite{Horst1996,Pinter1996,Strongin2000}.
+	\item локальные оценки константы $L_i$ в различных подобластях $D_i$ области поиска $D$ \cite{Kvasov2003,Sergeyev2010,Sergeyev2016}.
+	\item выбор оценок константы $L$ из некоторого множества возможных значений \cite{Gablonsky2001,Jones1993,Jones2009,Sergeyev2006}.
 \end{itemize}
@@
 
-В данной работе рассмотрен алгоритм, котором предлагается использовать две глобальных оценки константы Липшица, одна из которых значительно больше другой. 
+В данной работе рассмотрен алгоритм, котором предлагается использовать две глобальных оценки Lipschitz constant, одна из которых значительно больше другой. 
 The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
@@
 \section{Global search algorithm and dimensionality reduction}
-
+%Russian - два абзаца
 Типичным способом конструирования алгоритмов глобальной оптимизации является адаптация эффективных алгоритмов для решения одномерных задач к решению многомерных задач, see, for example, the diagonal partitions method in \cite{Sergeyev2006} or the simplicial partitions method in \cite{Zilinskas2008}.
 
-В данной статье используется подход, основанный на идеях редукции размерности с помощью Peano-Hilbert curves \cite{Strongin2000,Sergeyev2013}, which continuously and unambiguously maps the unit interval $[0,1]$ onto the $n$-dimensional cube $D$ from (\ref{D}).
+В данной статье используется подход, основанный на идеях редукции размерности с помощью Peano-Hilbert curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously maps the unit interval $[0,1]$ onto the $n$-dimensional cube $D$ from (\ref{D}).
 By using this kind of mapping it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem
@@
 \begin{equation}\label{R}
-R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r\mu},
+R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^*}{r\mu},
 \end{equation}
@@
 The theory of convergence of this algorithm is provided in \cite{Strongin2000}.
-%The modifications taking into account .... are given in [....].
 
@@
 
-Алгоритм глобального поиска, изложенной в предыдущем разделе, предназначен для решения многоэкстремальных задач, в которых целевая функция удовлетворяет условию Липшица.
+%Russian
+Алгоритм глобального поиска, изложенной в предыдущем разделе, предназначен для решения многоэкстремальных задач, в которых целевая функция удовлетворяет Lipschitz condition.
 При этом для для работы алгоритма не требуется задания  значений константы.
 Оценка константы осуществляется в процессе глобального поиска на основе имеющейся поисковой информации.
-В соответствии с теоремой \cite{Strongin2000}, последовательность точек испытаний $\{y^k\}$ будет сходиться к global minimizer $y^*$, если выполнено условие  
+В соответствии с теоремой из \cite{Strongin2000}, последовательность точек испытаний $\{y^k\}$ будет сходиться к global minimizer $y^*$, если выполнено условие  
 \begin{equation}\label{cond}
@@
 \end{equation}
-Таким образом, подходящий выбор параметра $r$ из (\ref{R}) позволяет использовать значение $r\mu / 2^{3-1/N}\sqrt{N+3}$ как оценку константы Липшица для целевой функции $\varphi(y)$.
+Таким образом, подходящий выбор параметра $r$ из (\ref{R}) позволяет использовать значение $r\mu / 2^{3-1/N}\sqrt{N+3}$ как оценку Lipschitz constant для целевой функции $\varphi(y)$.
 
 Выполнение условия (\ref{cond}) будет гарантировано при выборе большого значения параметра $r$, однако при этом метод выполнит большое количество испытаний до выполнения условия остановки.
-Выбор малого значения параметра $r$ (что соответствует заниженной оценке константы Липшица) значительно сократит число испытаний, но может нарушить сходимость к глобальному экстремуму.
+Выбор малого значения параметра $r$ (что соответствует заниженной оценке Lipschitz constant) значительно сократит число испытаний, но может нарушить сходимость к глобальному экстремуму.
 
-Перспективным является подход, при котором в правилах алгоритма будут использованы две оценки константы Липшица, одна из которых будет значительно выше другой, что соответствует использованию в алгоритме двух парметров $r_{glob}$ и $r_{loc}$, где $r_{glob} > r_{loc}$.
+Перспективным является подход, при котором в правилах алгоритма будут использованы две оценки Lipschitz constant, одна из которых будет значительно выше другой, что соответствует использованию в алгоритме двух парметров $r_{glob}$ и $r_{loc}$, где $r_{glob} > r_{loc}$.
 
-Правила алгоритма с двойной оценкой константы Липшица будут полностью повторять алгоритм глобального поиска, за исключением правила 3 (вычисление характеристики).
+Правила алгоритма с двойной оценкой Lipschitz constant будут полностью повторять алгоритм глобального поиска, за исключением правила 3 (вычисление характеристики).
 
@@
 \item
-Вычислить значение $R_{glob}(i)$, соответствующее большей оценке константы Липшица
+Вычислить значение $R_{glob}(i)$, соответствующее большей оценке Lipschitz constant
 \[
-R_{glob}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{glob}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r_{glob}\mu}.
+R_{glob}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{glob}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^*}{r_{glob}\mu}.
 \]
 \item
-Вычислить значение $R_{loc}(i)$, соответствующее меньшей оценке константы Липшица
+Вычислить значение $R_{loc}(i)$, соответствующее меньшей оценке Lipschitz constant
 \[
-R_{loc}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{loc}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r_{loc}\mu}.
+R_{loc}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{loc}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^*}{r_{loc}\mu}.
 \]
@@
 и зафиксировать значение $r = r_{loc}$, если  $\rho R_{loc}(i) > R_{glob}(i)$, иначе зафиксировать $r=r_{glob}$.
-Данное значение $r$ будет использовано в правиле 5 алгоритма при вычислении точки нового испытания.   
+\item
+Данное значение $r$ использовать в правиле 5 алгоритма при вычислении точки нового испытания.   
 \end{itemize}.
@@
 This generator of multiextremal functions is often used for the investigations of the global 
-optimization algorithms ~\cite{Paulavicius2014,Sergeyev2015,%Lebedev2015,Gergel2015,
-Barkalov2018}.
-%This generator allows multiextremal optimization problems to be generated with known properties (the number of localminima, the size of their domains of attraction, the global minimizer, etc.). 
+optimization algorithms ~\cite{Barkalov2018,Paulavicius2014,Sergeyev2015}.
 Six GKLS classes of differentiable test functions of the dimensions $N = 3,4,5$
@@
 algorithm generated a trial point $y^k$ in the vicinity of the global minimum,
-i.e. $\left|y^k-y^\ast\right| <\delta\left\|b-a\right\|$, where $\delta = 0.01$,
+i.e. $\left|y^k-y^\ast\right| <\delta\left\|b-a\right\|$, 
 $a$ and $b$ are the boundaries of the search domain $D$.
 
-%УКАЗАТЬ ПРАВИЛЬНЫЕ ПАРАМЕТРЫ
-When using the Global Search Algorithm, the parameter $r$ was set to $5$, 
-the evolvent construction parameter was fixed as $m = 10$. 
+%Russian
+Каждая серия задач была решена с помощью исходного алгоритма глобального поиска (GSA) и спомощью метода с двойной оценкой Lipschitz constant (GSA-DL).
+Для редукции размерности в обоих алгоритмах использовалась evolvent, построенная с параметром $m = 10$.
+Относительная точность поиска решения при $N=3,4$ составляла $\delta = 0.01$, а при $N=5$ -- $\delta = 0.03$.
 The maximum allowable number of iterations per problem was $K_{max} = 10^6$.
 
-Каждая серия задач была решена с помощью исходного алгоритма глобального поиска (GSA) и спомощью метода с двойной оценкой константы Липшица (GSA-DL).
-Среднее число итераций, выполненное алгоритмами, отражено в табл.
-\ref{tab:1}.
-При этом при решении задач класса Simple в GSA использовалось единственное значение параметра $r=4.8$, при решении задач класса Hard -- $r=5.6$.
+Среднее число итераций, выполненное алгоритмами, отражено в Table \ref{tab:1}.
+При этом при решении задач класса Simple в GSA использовалось значение параметра $r=4.8$, при решении задач класса Hard -- $r=5.6$.
 Указанное значение является минимальным (с точностью 0.1) значением, при котором успешно решились все задачи.
-При решении данных классов задач с использованием GSA-DL указанное ранее значение параметра $r$ выбиралось для верхней оценки константы Липшица, т.е.
-задавалось значение $r_{glob} = r$, которое дополнялось значениями $r_{loc}=1.8$ и $r_{loc}=2.4$.
-%Структуру таблицы можно взять из статьи
+При решении данных классов задач с использованием GSA-DL указанное ранее значение параметра $r$ выбиралось для верхней оценки Lipschitz constant, т.е.
+задавалось значение $r_{glob} = r$, которое дополнялось значениями $r_{loc}=1.8$, $r_{loc}=2.1$ $r_{loc}=2.4$.
 
@@
 		\hline\noalign{\smallskip}
-		$r$ & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
+		 & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
 		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \noalign{\smallskip}
@@
 		\noalign{\smallskip} \hline \noalign{\smallskip}
-									AGS	&	2502	&	2502	&	&	28254	&	28254	&	&	87261	&	87261	\\
-AGS-DL, $r_{loc}=2.4$	&	2502	&	2496	&	&	28254	&	27206	&	&	87261	&	85994	\\
-AGS-DL, $r_{loc}=1.8$	&	2502	&	2430	&	&	28254	&	30121	&	&	87261	&	78064	\\
+									AGS	&	2444	&	5345	& &	28415	&	77470	& &		&		\\
+AGS-DL, $r_{loc}=1.8$	&	1372	&	2632	& &	13273	&	37715	& &		&		\\
+AGS-DL, $r_{loc}=2.1$	&	1502	&	2805	& &	14826	&	38843	& &		&		\\
+AGS-DL, $r_{loc}=2.4$	&	1567	&	2868	& &	19447	&	40342	& &		&		\\
 		\noalign{\smallskip}\hline
@@
 
-Превосходство алгоритма GSA-DL над его прототипом наглядно подтверждается также и операционными характеристиками алгоритмов.
+Превосходство алгоритма GSA-DL над его прототипом  подтверждается также и операционными характеристиками алгоритмов.
 
-Результат решения всех задач серии будем представлять функцией $p(k)$, характеризующей долю задач, решенных за $k$ итераций.
-Такую функцию будем называть \textit{operational characteristic} of the algorithm..
+Пусть решается некоторая серия тестовых задач.
+Результат решения серии можно представить функцией $p(k)$, характеризующей долю задач, решенных за $k$ итераций.
+Такую функцию будем называть \textit{operational characteristic} of the algorithm.
 
-Операционные характеристики для методов GSA и GSA-DL, полученные при решении серий задач \textit{Simple} and \textit{Hard} размерностей $N=4$ и $N=5$ представлены соответственно на рис.
+Операционные характеристики для методов GSA и GSA-DL, полученные при решении серий задач \textit{Simple} and \textit{Hard} размерностей $N=4$ и $N=5$ представлены соответственно на Fig.
 \ref{oper4} и \ref{oper5}.
+Значения параметров $r$, используемых для оценки Lipschitz constant, приведены на рисунках.
 
@@
 \begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklss_4.png} \\ (a)}
+\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklss_5.png} \\ (a)}
 \end{minipage}
@@
 \begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklsh_4.png} \\ (b)}
+\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklsh_5.png} \\ (b)}
 \end{minipage}
@@
 
-Нижние кривые на обеих фигурах характеризуют GSA, а верхние линии — GSA-DL.
-Указанное расположение кривых показывает, что алгоритм с двумя оценками константы Липшица обеспечивает в среднем значительно более быстрое решение задач данного класса, чем алгоритм с использованием единственной оценки константы Липшица.
-
-\section{Conclusion}
+Нижние кривые на Fig.
+\ref{oper4} и \ref{oper5} характеризуют GSA, а верхние линии — GSA-DL.
+Указанное расположение кривых показывает, что алгоритм с двумя оценками Lipschitz constant обеспечивает в среднем значительно более быстрое решение задач серии, чем алгоритм с использованием единственной оценки константы.
 

### 2019-05-20 2019/NUMTA_Barkalov/LNCS/numta19.tex
after translation

@@
 %Russian
-Предположение липшицевости целевой функции является типичным для многих подходов к разработке алгоритмов глобальной оптимизации \cite{Evtushenko2013,Zilinskas2010,Pinter1996,Strongin2000}.
-При этом наиболее важной проблемой, так или иначе решаемой в данных алгоритмах, является адаптивная оценка неизвестной Lipschitz constant на основе полученной поисковой информации.
+The assumption of the objective function to be Lipschitzian is typical for many approaches to the development of the global optimization algorithms \cite{Evtushenko2013,Zilinskas2010,Pinter1996,Strongin2000}.
+At that, the adaptive estimate of the unknown Lipschitz constant based on the obtained search information is the most important problem being solved in these algorithms anyway.
+The Lipschitz constant value affects the convergence rate of the global optimization algorithms essentially. Therefore, the issue of its correct estimate is so important. 
+The underestimate of the constant true value may result in losing the convergence of the algorithm to the global solution.
+At the same time, too large value of the estimate of the constant for the objective function not matching its real behavior results in slow convergence of the algorithm to the global minimizer.
 
-Значение Lipschitz constant существенно влияет на скорость сходимости алгоритмов липшицевой глобальной оптимизации, поэтому столь важным является вопрос ее корректной оценки.
-Заниженная оценка инстинного значения константы может привести к потере сходимости алгоритма к глобальному решению.
-В то же время слишком большое значение оценки константы $L$ для целевой функции, не соответствующее ее истинному поведению, влечет за собой медленную сходимость алгоритма к точке глобального минимума.
-
-Известны несколько типичных способов адаптивного оценивания Lipschitz constant:
+Several typical methods of adaptive estimation of the Lipschitz constant are known:
 \begin{itemize}
-	\item глобальная оценка константы $L$ во всей области поиска $D$ \cite{Horst1996,Pinter1996,Strongin2000}.
-	\item локальные оценки константы $L_i$ в различных подобластях $D_i$ области поиска $D$ \cite{Kvasov2003,Sergeyev2010,Sergeyev2016}.
-	\item выбор оценок константы $L$ из некоторого множества возможных значений \cite{Gablonsky2001,Jones1993,Jones2009,Sergeyev2006}.
+	\item global estimation of the constant $L$ in the whole search domain $D$ \cite{Horst1996,Pinter1996,Strongin2000}.
+	\item local estimations of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ \cite{Kvasov2003,Sergeyev2010,Sergeyev2016}.
+	\item the choice of the estimates of the constant $L$ from a set of possible values \cite{Gablonsky2001,Jones1993,Jones2009,Sergeyev2006}.
 \end{itemize}
 
-Каждый из указанных подходов обладает своими достоинствами и недостатками.
-Например, использование только глобальной оценки во всей области поиска может замедлить сходимость алгоритма к точке глобального минимума.
-А использование локальных оценок, ускоряющих сходимость, требует адекватной настройки параметров алгоритма для сохранения глобальной сходимости.
+Each approach from the above has its own advantages and disadvantages.
+For example, the use of the global estimate over the whole search domain can slow down the convergence of the algorithm to the global minimizer.
+The use of the local estimates accelerating the convergence requires an adequate adjustment of the algorithm parameters in order to preserve the global convergence.
 
-В данной работе рассмотрен алгоритм, котором предлагается использовать две глобальных оценки Lipschitz constant, одна из которых значительно больше другой. 
+In the present work, an algorithm, in which it is proposed to use two global estimates of the Lipschitz constant, one being much greater than another, is considered. 
 The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
-Выбор того, какая из двух оценок будет использоваться в правилах алгоритма, осуществляется адаптивно, в зависимости от поведения функции и фазы поиска.
+The choice of the one of two estimates to be used in the algorithm rules is performed adaptively subject to the function behavior and to the search phase.
 
-Строгое обоснование предложенного подхода выходит за рамки данной первоначальной публикации и будет сделано в последующих работах.
-Здесь же приведем результаты вычислительных экспериментов, наглядно демонстрирующие эффективность нового алгоритма.
-При проведении вычислительных экспериментов было решено несколько сотен тестовых многоэкстремальных задач разной размерности.
+The rigorous substantiation of the proposed approach goes beyond the present initial publication and will be done in the forthcoming works.
+Here we present the results of numerical experiments clearly demonstrating the efficiency of the novel algorithm.
+When conducting the numerical experiments, several hundred multiextremal test problems of various dimensionalities have been solved.
 
 \section{Global search algorithm and dimensionality reduction}
-%Russian - два абзаца
-Типичным способом конструирования алгоритмов глобальной оптимизации является адаптация эффективных алгоритмов для решения одномерных задач к решению многомерных задач, see, for example, the diagonal partitions method in \cite{Sergeyev2006} or the simplicial partitions method in \cite{Zilinskas2008}.
+%Russian
 
-В данной статье используется подход, основанный на идеях редукции размерности с помощью Peano-Hilbert curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously maps the unit interval $[0,1]$ onto the $n$-dimensional cube $D$ from (\ref{D}).
-By using this kind of mapping it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem
+The adaptation of the efficient algorithms of solving the one-dimensional problems to solving the multidimensional problems is typical method of constructing the global optimization algorithms, see, for example, the diagonal partitions method in \cite{Sergeyev2006} or the simplicial partitions method in \cite{Zilinskas2008}.
+
+In the present paper, the approach based on the ideas of dimensionality reduction with the use of the Peano-Hilbert curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously maps the unit interval $[0,1]$ onto the $n$-dimensional cube $D$ from (\ref{D}) has been used.
+By using this kind of mapping, it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem
 \[
@@
 the points $y^0=y(0), y^1=y(1)$. The choice of the point $y^{k+1},k\geq 1,$  
-for the next $(k+1)$-th trial is defined by the following rules.
+for the next $(k+1)^{\rm th}$ trial is defined by the following rules.
 
@@
 %Russian
-Алгоритм глобального поиска, изложенной в предыдущем разделе, предназначен для решения многоэкстремальных задач, в которых целевая функция удовлетворяет Lipschitz condition.
-При этом для для работы алгоритма не требуется задания  значений константы.
-Оценка константы осуществляется в процессе глобального поиска на основе имеющейся поисковой информации.
-В соответствии с теоремой из \cite{Strongin2000}, последовательность точек испытаний $\{y^k\}$ будет сходиться к global minimizer $y^*$, если выполнено условие  
+The global search algorithm presented in previous section is intended for solving the multiextremal problems, in which the objective function satisfies the Lipschitz condition.
+It is not necessary to define the value of the constant for the algorithm convergence.
+The estimation of the constant is performed in the course of global search on the base of available search information.
+According to the theorem in \cite{Strongin2000}, the sequence of the trial points $\{y^k\}$ will converge to the global minimizer $y^*$ if the condition 
 \begin{equation}\label{cond}
@@
 \end{equation}
-Таким образом, подходящий выбор параметра $r$ из (\ref{R}) позволяет использовать значение $r\mu / 2^{3-1/N}\sqrt{N+3}$ как оценку Lipschitz constant для целевой функции $\varphi(y)$.
+\noindent is satisfied.
+Thus,  an appropriate choice of the parameter $r$ from (\ref{R}) allows using the value $r\mu / 2^{3-1/N}\sqrt{N+3}$ as an estimate of the Lipschitz constant for the objective function $\varphi(y)$.
 
-Выполнение условия (\ref{cond}) будет гарантировано при выборе большого значения параметра $r$, однако при этом метод выполнит большое количество испытаний до выполнения условия остановки.
-Выбор малого значения параметра $r$ (что соответствует заниженной оценке Lipschitz constant) значительно сократит число испытаний, но может нарушить сходимость к глобальному экстремуму.
+Satisfying the condition (\ref{cond}) will be guaranteed at choosing large enough value of the parameter $r$.
+However, in this case the method will perform a large number of trials until the stop condition is satisfied.
+The choice of small value of the parameter $r$ (that corresponds to the lower estimate of the Lipschitz constant) would considerably reduce the number of trials but may violate the convergence to the global extremum.
 
-Перспективным является подход, при котором в правилах алгоритма будут использованы две оценки Lipschitz constant, одна из которых будет значительно выше другой, что соответствует использованию в алгоритме двух парметров $r_{glob}$ и $r_{loc}$, где $r_{glob} > r_{loc}$.
+An approach, in which two estimates of the Lipschitz constant %, one of which is essentially greater than another, 
+are used in the rules of algorithm, is a promising one. 
+This approach implies the use of two parameters $r_{glob}$ and $r_{loc}$ where $r_{glob} > r_{loc}$ in the algorithm.
 
-Правила алгоритма с двойной оценкой Lipschitz constant будут полностью повторять алгоритм глобального поиска, за исключением правила 3 (вычисление характеристики).
+The rules of the algorithm with two estimates of the Lipschitz constant reproduce the ones of global search algorithm completely except Rule 3 (computing the characteristic).
 
-Новое правило вычисления характеристики $R(i)$ интервала $(x_{i-1}, x_i)$ будет состоять из следующих действий:
+New rule of calculating the characteristic $R(i)$ of the interval $(x_{i-1}, x_i)$ will consist of the following operations:
 \begin{itemize}
 \item
-Вычислить значение $R_{glob}(i)$, соответствующее большей оценке Lipschitz constant
+Calculate the value $R_{glob}(i)$ corresponding to the larger estimate of the Lipschitz constant
 \[
@@
 \item
-Вычислить значение $R_{loc}(i)$, соответствующее меньшей оценке Lipschitz constant
+Calculate the value $R_{loc}(i)$ corresponding to the smaller estimate of the Lipschitz constant
 \[
@@
 \item
-Определить характеристику интервала $R(i)$ как
+Determine the characteristic $R(i)$ as
 \begin{equation}\label{pho}
-R(i) = \max\{\rho R_{loc}(i),R_{glob}(i)\}, где \rho = \left(\frac{1-1/r_{glob}}{1-1/r_{loc}}\right)^2,
+R(i) = \max\{\rho R_{loc}(i),R_{glob}(i)\}, where \rho = \left(\frac{1-1/r_{glob}}{1-1/r_{loc}}\right)^2,
 \end{equation}
-и зафиксировать значение $r = r_{loc}$, если  $\rho R_{loc}(i) > R_{glob}(i)$, иначе зафиксировать $r=r_{glob}$.
+and fix the value $r = r_{loc}$ if $\rho R_{loc}(i) > R_{glob}(i)$, otherwise fix $r=r_{glob}$.
 \item
-Данное значение $r$ использовать в правиле 5 алгоритма при вычислении точки нового испытания.   
+This value of $r$ is used in Rule 5 of the algorithm in the computing of the next trial point.   
 \end{itemize}.
 
-Указанный способ вычисления характеристики имеет следующее обоснование. 
-Пусть текущее минимальное значение $z^*$ из (\ref{z}) достигается в левой точке $i$-го интервала, т.е. $z^* = z_{i-1}$. Тогда, в соответствии с правилом (\ref{R}), будут верно неравенство
+This method of computing the characteristic is substantiated as follows. 
+Let current minimum value of $z^*$ from (\ref{z}) is achieved at the left point of the $i^{\rm th}$ interval, i.e.
+$z^* = z_{i-1}$.
+Then, according to the rule (\ref{R}) the following inequality will be true:
 \[
@@
 \]
-Следовательно, две нижние оценки для двух характеристик текущего лучшего интервала $R_{loc}(i)$ и $R_{glob}(i)$, вычисленных с разными параметрами $r_{loc}$ и $r_{glob}$, будут иметь разные значения.
-Однако если первую из них домножить на коэффициент $\rho$ из (\ref{pho}), то нижние оценки характеристик лучшего интервала будут совпадать, что сделает сравнимыми и характеристики, вычисленные для остальных интервалов.
+Consequently, two lower estimates of the two characteristics of current best interval $R_{loc}(i)$ and $R_{glob}(i)$ calculated with different parameters $r_{loc}$ and $r_{glob}$ will take different values.
+However, if the former is multiplied by the coefficient $\rho$ from (\ref{pho}), the lower estimates of the characteristics of the best interval would match each other that would make the characteristics calculated for other intervals comparable as well.
 
@@
 %Russian
-Каждая серия задач была решена с помощью исходного алгоритма глобального поиска (GSA) и спомощью метода с двойной оценкой Lipschitz constant (GSA-DL).
-Для редукции размерности в обоих алгоритмах использовалась evolvent, построенная с параметром $m = 10$.
-Относительная точность поиска решения при $N=3,4$ составляла $\delta = 0.01$, а при $N=5$ -- $\delta = 0.03$.
+Each series of problems has been solved by the original global search algorithm (GSA) and by the method with two estimates of the Lipschitz constant (GSA-DL).
+The evolvent constructed using the parameter $m = 10$ was used for the dimensionality reduction in both algorithms.
+The relative accuracy of the solution search was $\delta = 0.01$ for $N=3,4$ and $\delta = 0.03$ for $N=5$.
 The maximum allowable number of iterations per problem was $K_{max} = 10^6$.
 
-Среднее число итераций, выполненное алгоритмами, отражено в Table \ref{tab:1}.
-При этом при решении задач класса Simple в GSA использовалось значение параметра $r=4.8$, при решении задач класса Hard -- $r=5.6$.
-Указанное значение является минимальным (с точностью 0.1) значением, при котором успешно решились все задачи.
-При решении данных классов задач с использованием GSA-DL указанное ранее значение параметра $r$ выбиралось для верхней оценки Lipschitz constant, т.е.
-задавалось значение $r_{glob} = r$, которое дополнялось значениями $r_{loc}=1.8$, $r_{loc}=2.1$ $r_{loc}=2.4$.
+The averaged numbers of iterations performed by the algorithms are presented in Table \ref{tab:1}.
+At that, for GSA the values of parameter $r=4.8$ when solving the problems of Simple class and $r=5.6$ when solving the problems of Hard class were used. 
+These values are the minimum ones (with the accuracy 0.1), at which all problems have been solved successfully.
+When solving the problems from the above classes by GSA-DL, the value of parameter $r$ specified above was selected for the upper estimate of the Lipschitz constant, i.e.
+the value $r_{glob} = r$ was set, which was complemented by the values $r_{loc}=1.8$, $r_{loc}=2.1$ and $r_{loc}=2.4$.
 
@@
 
-Превосходство алгоритма GSA-DL над его прототипом  подтверждается также и операционными характеристиками алгоритмов.
+The advantage of the GSA-DL algorithm over its prototype is confirmed by the operational characteristics of the algorithms as well.
 
-Пусть решается некоторая серия тестовых задач.
-Результат решения серии можно представить функцией $p(k)$, характеризующей долю задач, решенных за $k$ итераций.
-Такую функцию будем называть \textit{operational characteristic} of the algorithm.
+Assume a series of test problems to be solved.
+The results of solving the series can be presented by a function $p(k)$ featuring the fraction of the total number of problems solved in $k$ iterations.
+Such a function will be called the \textit{operational characteristic} of the algorithm.
 
-Операционные характеристики для методов GSA и GSA-DL, полученные при решении серий задач \textit{Simple} and \textit{Hard} размерностей $N=4$ и $N=5$ представлены соответственно на Fig.
-\ref{oper4} и \ref{oper5}.
-Значения параметров $r$, используемых для оценки Lipschitz constant, приведены на рисунках.
+The operational characteristics for the GSA and GSA-DL methods obtained when solving the \textit{Simple} and \textit{Hard} problem series with the dimensionalities $N=4$ and $N=5$ are presented in Fig.
+\ref{oper4} and \ref{oper5}, respectively.
+The values of the parameters $r$ used for estimating the Lipschitz constant are given in the figures.
 
@@
 
-Нижние кривые на Fig.
-\ref{oper4} и \ref{oper5} характеризуют GSA, а верхние линии — GSA-DL.
-Указанное расположение кривых показывает, что алгоритм с двумя оценками Lipschitz constant обеспечивает в среднем значительно более быстрое решение задач серии, чем алгоритм с использованием единственной оценки константы.
+The lower curves in Fig.
+\ref{oper4} and \ref{oper5} feature GSA whereas the upper ones --- GSA-DL.
+Such positions of the curves show the algorithm with two estimates of the Lipschitz constant provides much faster solving of the problem series in average that the algorithm using single estimate of the constant.
 
-%
 % ---- Bibliography ----

### 2019-05-22 2019/NUMTA_Barkalov/LNCS/numta19.tex

@@
 %
-\title{Acceleration of Global Search through Dual Lipschitz Constant Estimates\thanks{This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.}}
+\title{Acceleration of Global Search by Implementing Dual Estimates for Lipschitz Constant\thanks{This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.}}
 %
@@
 The theory of convergence of this algorithm is provided in \cite{Strongin2000}. 
+The algorithm is allows for an efficient
+parallelization for shared and distributed memory \cite{Gergel2003}  and for accelerators \cite{Gergel2016}.
 
@@
 The estimation of the constant is performed in the course of global search on the base of available search information.
-According to the theorem in \cite{Strongin2000}, the sequence of the trial points $\{y^k\}$ will converge to the global minimizer $y^*$ if the condition 
+According to the theorem from \cite{Strongin2000}, the sequence of the trial points $\{y^k\}$ will converge to the global minimizer $y^*$ if the condition 
 \begin{equation}\label{cond}
@@
 \begin{equation}\label{pho}
-R(i) = \max\{\rho R_{loc}(i),R_{glob}(i)\}, where \rho = \left(\frac{1-1/r_{glob}}{1-1/r_{loc}}\right)^2,
+R(i) = \max\{\rho R_{loc}(i),R_{glob}(i)\}, \textrm{where} \; \rho = \left(\frac{1-1/r_{glob}}{1-1/r_{loc}}\right)^2,
 \end{equation}
@@
 The evolvent constructed using the parameter $m = 10$ was used for the dimensionality reduction in both algorithms.
-The relative accuracy of the solution search was $\delta = 0.01$ for $N=3,4$ and $\delta = 0.03$ for $N=5$.
+The relative accuracy of the solution search was $\delta = 0.01$.% for $N=3,4$ and $\delta = 0.02$ for $N=5$.
 The maximum allowable number of iterations per problem was $K_{max} = 10^6$.
@@
 the value $r_{glob} = r$ was set, which was complemented by the values $r_{loc}=1.8$, $r_{loc}=2.1$ and $r_{loc}=2.4$.
+The number of unsolved problems is specified in brackets.
 
@@
 		\noalign{\smallskip} \hline \noalign{\smallskip}
-									AGS	&	2444	&	5345	& &	28415	&	77470	& &		&		\\
-AGS-DL, $r_{loc}=1.8$	&	1372	&	2632	& &	13273	&	37715	& &		&		\\
-AGS-DL, $r_{loc}=2.1$	&	1502	&	2805	& &	14826	&	38843	& &		&		\\
-AGS-DL, $r_{loc}=2.4$	&	1567	&	2868	& &	19447	&	40342	& &		&		\\
+									AGS	&	2444	&	5345	& &	28415	&	77470	& & 25220(1) & 126138(4)		\\
+AGS-DL, $r_{loc}=1.8$	&	1372	&	2632	& &	13273	&	37715	& &	12702    & 94296(1)			\\
+AGS-DL, $r_{loc}=2.1$	&	1502	&	2805	& &	14826	&	38843	& &	15213    & 90792(2)			\\
+AGS-DL, $r_{loc}=2.4$	&	1567	&	2868	& &	19447	&	40342	& &	18239    & 106438(2)		\\
 		\noalign{\smallskip}\hline

### 2019-06-07 2019/RSD_Integer/RSD2019.tex
After review

@@
 
-\title{Parallel Global Optimization for Non-Convex Mixed-Integer Problems
-\thanks{This study was supported by the Russian Science Foundation, 
-project No.\,16-11-10150.}}
+\title{Parallel Global Optimization for Non-Convex Mixed-Integer Problems\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}}
 \author{Konstantin Barkalov \and Ilya Lebedev %\Letter 
@@
 A detailed description of the algorithm convergence theory is presented in \cite{Strongin2000}.
+The results of comparison of the algorithm with other sequential and parallel global optimization algorithms have been presented in \cite{Sovrasov2019}.
 
@@
 
+\bibitem{Sovrasov2019}
+Sovrasov, V., Bevzuk, S.: Comparison of several sequential and parallel derivative-free global optimization algorithms. 	Communications in Computer and Information Science (2019) (article in press)
+
 \bibitem{Barkalov2016} 

### 2019-06-26 2019/MOTOR_Sovrasov/slides/slides.tex
MOTOR: updae slides

@@
       \item Usually store and use the whole history of trials accumulated during search;
-      \item Require non-trivial parallelization schemes;
       \item Under some assumptions convergence to the global solution is guaranteed.
@@
       \item Require constant amount of memory to store internal state of some random process or individuals of population;
-      \item In most cases allows wide-scale parallelization by trials or by running many copies of the same method
-      with different seeds in parallel;
       \item Convergence is guaranteed in probabilistic sense only.
@@
     \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Test problems}
-  \begin{columns}
-    \begin{column}{0.5\textwidth}
-      Generator GKLS was employed to construct 8 sets of 100 test problems:
-      \begin{displaymath}
-        \begin{matrix}
-          f(x)=
-          \left\{
-          \begin{matrix}
-          C_i(x), x \in S_i, i\in 2,\dots ,m \\
-          \Vert x-T \Vert^2 + t, x\not\in S_2,\dots,S_m
-          \end{matrix} \right.
-        \end{matrix}
-      \end{displaymath}
-
-      The generator allows to adjust:
-      \begin{itemize}
-        \item the number of local minimums;
-        \item the size of the global minima attraction region;
-        \item the space dimension.
-      \end{itemize}
-    \end{column}
-    \begin{column}{0.5\textwidth}
-      \centerline{\includegraphics[width=0.9\textwidth]{gkls_color.png}}
-    \end{column}
-  \end{columns}
-  Also one set of 100 pre-defined highly multiextremal functions (with 10-30 local minimums) was used.
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results of sequential methods}
-
-  \begin{table}
-  \begin{center}
-  \caption{Averaged number of trials executed by sequential methods for solving the test
-  optimization problems}
-  \resizebox{\textwidth}{!}{%
-    \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
-      \hline
-      & AGS & AGS\(l\) & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
-    \hline
-    \(F_{GR}\)     & 193.1 &  \textbf{158.3} & 400.3 & 182.3 & 214.9 & 947.2 & 691.2 & 1257.3 &
-  1336.8 \\
-    \hline
-    GKLS 2d Simple &  254.9 & 217.6 & 510.6 & \textbf{189.0} & 255.2 & 556.8 & 356.3 & 952.2
-  & 1251.5 \\
-    \hline
-    GKLS 2d Hard   &  728.7 & \textbf{488.0} & 844.7 & 985.4 & 1126.7 & 1042.5 & 1637.9 &
-  1041.1 & 2532.2 \\
-    \hline
-    GKLS 3d Simple &  1372.1 & 1195.3 & 4145.8 & \textbf{973.6} & 1477.8 & 4609.2 & 2706.5 &
-  5956.94 & 3856.1 \\
-    \hline
-    GKLS 3d Hard   &  3636.1 & \textbf{1930.5} & 6787.0 & 2298.7 & 3553.3 & 5640.1 & 4708.4 &
-  6914.3 & 7843.2 \\
-    \hline
-    GKLS 4d Simple &  26654.1 & 11095.7 & 37436.8 & \textbf{7824.3} & 15994.1 & 41514.3 &
-  21417.9 & 19157.7 & 59895.4 \\
-    \hline
-    GKLS 4d Hard   &  54536.8 &  \textbf{23167.8} &  73779.3 &  23204.4 &  54489.9 &  80247.2
-  &  68815.5 &  27466.1 &  109328.1  \\
-    \hline
-    GKLS 5d Simple &  29810.0 & 11529.0 & 143575.0 & \textbf{7166.5} & 13970.5 & 52647.6 &
-  34255.3 & 73074.5 & 91580.4 \\
-    \hline
-    GKLS 5d Hard   &  113129.1 & 67652.7 & 165192.8 & \textbf{66327.4} & 164390.6 & 138766.2
-  & 116973.1 & 105496.9 & 155123.8 \\
-    \hline
-  \end{tabular}}
-    \label{tab:trials}
-  \end{center}
-  \end{table}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results of sequential methods}
-  \begin{table}
-  \begin{center}
-  \caption{Number of test optimization problems solved by sequential methods}
-    \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
-      \hline
-      & AGS & AGS\(l\) & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
-    \hline
-    \(F_{GR}\)     &  100 & 100 & 76  & 100 & 100 & 97  & 96  & 96  & 67\\
-    \hline
-    GKLS 2d Simple &  100 & 100 & 85  & 100 & 100 & 100 & 100 & 98  & 90\\
-    \hline
-    GKLS 2d Hard   &  100 & 100 & 74  & 100 & 100 & 100 & 93  & 85  & 77 \\
-    \hline
-    GKLS 3d Simple &  100 & 97  & 75  & 100 & 100 & 100 & 89  & 86  & 44 \\
-    \hline
-    GKLS 3d Hard   &  100  & 99   & 72   & 100  & 99   & 100  & 88   & 77   & 43 \\
-    \hline
-    GKLS 4d Simple &  100 & 100 & 46  & 100 & 100 & 94  & 78  & 59  & 16 \\
-    \hline
-    GKLS 4d Hard   &  100 & 100 & 47  & 99  & 97  & 94  & 72  & 32  & 10  \\
-    \hline
-    GKLS 5d Simple &  100 & 100 & 68  & 100 & 100 & 98  & 100 & 77  & 9  \\
-    \hline
-    GKLS 5d Hard   &  97  & 99  & 42  & 100 & 90  & 79  & 84  & 48  & 8 \\
-    \hline
-    \end{tabular}
-    \label{tab:solved}
-  \end{center}
-  \end{table}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Parallel optimization methods}
-  \begin{itemize}
-    \item Parallel AGS (Globalizer system);
-    \item Parallel ACO (MIDACO system).
   \end{itemize}
@@
 \begin{frame}
+  \frametitle{Hyperparameters tuning in AGS}
+  AGS selects the best intervals according to the following characteristic:
+  \begin{displaymath}
+    R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{(rH)^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{rH},1<i<k+1
+  \end{displaymath}
+  where $L$ is current estimation on Holder constant, $r$ is reliability coefficient.
+  AGS is very sensitive to choice of $r$:
+  \begin{itemize}
+    \item low value can lead to fast convergence to a local minima
+    \item high value leads to very slow convergence
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Hyperparameters tuning in AGS}
+  In order to resolve the problem of choosing $r$ to some extent,
+  let us use the following scheme:
+  \begin{itemize}
+    \item execute $q$ iterations of AGS with $r=r_{max}$;
+    \item execute $q$ iterations of AGS with $r=r_{min}$;
+    \item repeat the above steps either until convergence or until the allowed number of iterations are
+  exhausted.
+\end{itemize}
+  Now we have 3 parameters instead of one, but the method is not so sensitive to them.
+  Let $q=50\cdot\log_2(N-1)\cdot N^2$, $r_{min}=3,\:r_{max}=2\cdot r_{min}$ in all the future experiments.
+\end{frame}
+
+\begin{frame}
+  \frametitle{Test problems}
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+      Generator GKLS was employed to construct 8 sets of 100 test problems:
+      \begin{displaymath}
+        \begin{matrix}
+          f(x)=
+          \left\{
+          \begin{matrix}
+          C_i(x), x \in S_i, i\in 2,\dots ,m \\
+          \Vert x-T \Vert^2 + t, x\not\in S_2,\dots,S_m
+          \end{matrix} \right.
+        \end{matrix}
+      \end{displaymath}
+
+      The generator allows to adjust:
+      \begin{itemize}
+        \item the number of local minimums;
+        \item the size of the global minima attraction region;
+        \item the space dimension.
+      \end{itemize}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \centerline{\includegraphics[width=0.9\textwidth]{gkls_color.png}}
+    \end{column}
+  \end{columns}
+  Also one set of 100 pre-defined highly multiextremal functions (with 10-30 local minimums) was used.
+\end{frame}
+
+\begin{frame}
+  \frametitle{Experimental setup}
+  \begin{table}
+  \begin{center}
+  \caption{Trials limits and relative precision for the test problem classes}
+    \begin{tabular}{|l|{c}|{c}|}
+      \hline
+    Problems class & Trials limit & $\alpha$\\
+    \hline
+    \(F_{GR}\) & 5000 & 0.01 \\
+    \hline
+    GKLS 2d Simple & 8000 & 0.01 \\
+    \hline
+    GKLS 2d Hard & 9000 & 0.01 \\
+    \hline
+    GKLS 3d Simple & 15000 & 0.01 \\
+    \hline
+    GKLS 3d Hard & 25000 & 0.01 \\
+    \hline
+    GKLS 4d Simple & 150000 & $\sqrt[4]{10^{-6}}$ \\
+    \hline
+    GKLS 4d Hard & 250000 & $\sqrt[4]{10^{-6}}$ \\
+    \hline
+    GKLS 5d Simple & 350000 & $\sqrt[5]{10^{-7}}$ \\
+    \hline
+    GKLS 5d Hard & 600000 & $\sqrt[5]{10^{-7}}$ \\
+    \hline
+    \end{tabular}
+    \label{tab:limits}
+  \end{center}
+  \end{table}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Experimental setup}
+  \begin{table}
+  \begin{center}
+  \caption{Class-specific parameters of the optimization algorithms}
+    \begin{tabular}{|l|{c}|{c}|{c}|}
+      \hline
+      & AGS & CRS & DE\\
+    \hline
+    \(F_{GR}\) & \(r=3\) & popsize=150 & mutation=(1.1,1.9), popsize=60 \\
+    \hline
+    GKLS 2d Simple & \(r=4.6\) & popsize=200 & mutation=(1.1,1.9), popsize=60 \\
+    \hline
+    GKLS 2d Hard & \(r=6.5\) & popsize=400 & mutation=(1.1,1.9), popsize=60 \\
+    \hline
+    GKLS 3d Simple & \(r=3.7\) & popsize=1000 & mutation=(1.1,1.9), popsize=70 \\
+    \hline
+    GKLS 3d Hard & \(r=4.4\) & popsize=2000 & mutation=(1.1,1.9), popsize=80 \\
+    \hline
+    GKLS 4d Simple & \(r=4.7\) & popsize=8000 & mutation=(1.1,1.9), popsize=90 \\
+    \hline
+    GKLS 4d Hard & \(r=4.9\) & popsize=16000 & mutation=(1.1,1.9), popsize=100 \\
+    \hline
+    GKLS 5d Simple & \(r=4\) & popsize=25000 & mutation=(1.1,1.9), popsize=120 \\
+    \hline
+    GKLS 5d Hard & \(r=4\) & popsize=30000 & mutation=(1.1,1.9), popsize=140 \\
+    \hline
+  \end{tabular}
+    \label{tab:params}
+  \end{center}
+  \end{table}
+  \begin{itemize}
+    \item in the DIRECT and DIRECT\(l\) methods, the parameter \(\epsilon=10^{-4}\);
+    \item in the SDA method, the parameter \(visit=2.72\).
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results of different methods}
+
+  \begin{table}
+  \begin{center}
+  \caption{Averaged number of trials executed by the methods for solving the test
+  optimization problems}
+  \resizebox{\textwidth}{!}{%
+    \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
+      \hline
+      & AGS & AGS-AR & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
+    \hline
+    \(F_{GR}\)     & 193.1 & 248.3 & 400.3 & \textbf{182.2} & 214.9 & 947.2 & 691.2 & 1257.3 & 1336.8 \\
+    \hline
+    GKLS 2d Simple & 254.9 & 221.6 & 510.6 & \textbf{189.0} & 255.2 & 556.8 & 356.3 & 952.2 & 1251.5 \\
+    \hline
+    GKLS 2d Hard   & \textbf{728.7} & 785.0 & 844.7 & 985.4 & 1126.7 & 1042.5 & 1637.9 & 1041.1 & 2532.2 \\
+    \hline
+    GKLS 3d Simple &  1372.1 & 1169.5 & 4145.8 & \textbf{973.6} & 1477.8 & 4609.2 & 2706.5 & 5956.9 & 3856.1 \\
+    \hline
+    GKLS 3d Hard   &  3636.1 & \textbf{1952.1} & 6787.0 & 2298.7 & 3553.3 & 5640.1 & 4708.4 & 6914.3 & 7843.2 \\
+    \hline
+    GKLS 4d Simple &  5729.8 & \textbf{4919.1} & 19883.6 & 7328.8 & 15010.0 & 41484.8 & 22066.0 & 6271.2 & 29359.2 \\
+    \hline
+    GKLS 4d Hard   &  13113.4 & \textbf{12860.1} & 27137.4 & 22884.4 & 55596.1 & 80220.1 & 68048.0 & 12487.6 & 58925.5  \\
+    \hline
+    GKLS 5d Simple &  \textbf{5821.5} & 6241.3 & 62921.7 & 5966.1 & 10795.5 & 52609.2 & 34208.8 & 20859.4 & 69206.8 \\
+    \hline
+    GKLS 5d Hard   &  \textbf{17008.6} & 21555.1 & 87563.9 & 61657.3 & 148637.8 & 138011.8 & 115634.6 & 26850.0 & 141886.5 \\
+    \hline
+  \end{tabular}}
+    \label{tab:trials}
+  \end{center}
+  \end{table}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results of different methods}
+  \begin{table}
+  \begin{center}
+  \caption{Number of test optimization problems solved by the methods}
+  \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
+    \hline
+    & AGS & AGS-AR & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
+  \hline
+  \(F_{GR}\)     &  100 & 100 & 76 & 100 & 100 & 97 & 96 & 96 & 67\\
+  \hline
+  GKLS 2d Simple &  100 & 100 & 85 & 100 & 100 & 100 & 100 & 98 & 90\\
+  \hline
+  GKLS 2d Hard   &  100 & 97 & 74 & 100 & 100 & 100 & 93 & 85 & 77 \\
+  \hline
+  GKLS 3d Simple &  100 & 100 & 75 & 100 & 100 & 100 & 89 & 86 & 44 \\
+  \hline
+  GKLS 3d Hard   &  100 & 100 & 72 & 100 & 99 & 100 & 88 & 77 & 43 \\
+  \hline
+  GKLS 4d Simple &  100 & 100 & 74 & 100 & 100 & 94 & 82 & 68 & 72 \\
+  \hline
+  GKLS 4d Hard   &  100 & 100 & 60 & 99 & 99 & 94 & 73 & 55 & 69  \\
+  \hline
+  GKLS 5d Simple &  100 & 100 & 86 & 100 & 100 & 98 & 100 & 88 & 82  \\
+  \hline
+  GKLS 5d Hard   &  100 & 100 & 77 & 100 & 93 & 79 & 86 & 77 & 78 \\
+  \hline
+  \end{tabular}
+    \label{tab:solved}
+  \end{center}
+  \end{table}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results of different methods}
+  \begin{figure}[ht]
+    \centering
+    \subfloat[4d Simple]{{\includegraphics[width=.5\textwidth]{gklss4d.pdf}}\label{fig:s4d}}
+    \subfloat[4d Hard]{{\includegraphics[width=.5\textwidth]{gklsh4d.pdf}}\label{fig:h4d}}
+  \end{figure}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results of different methods}
+  \begin{figure}[ht]
+    \centering
+    \subfloat[5d Simple]{{\includegraphics[width=.5\textwidth]{gklss5d.pdf}}\label{fig:s4d}}
+    \subfloat[5d Hard]{{\includegraphics[width=.5\textwidth]{gklsh5d.pdf}}\label{fig:h4d}}
+  \end{figure}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Robustness of the AGS-AR to hyperparemeters}
+  \begin{figure}[ht]
+    \centering
+    \includegraphics[width=.6\textwidth]{ar_stab.pdf}
+    \caption{Operating characteristics of AGS and AGS-AR with different hyperparameters
+    when solving problems from the GKLS 5d Simple classe.}
+    \label{fig:stability}
+  \end{figure}
+\end{frame}
+
+\begin{frame}
   \frametitle{Conclusions}
@@
     \begin{itemize}
-      \item AGS\(l\) method implemented in the Globalizer system has demonstrated the convergence
+      \item The proposed modification of the stock AGS, AGS-AR allows to pay less attention to initial hyperparameter tuning and
+      performs on-par with properly tuned AGS;
+      \item AGS-AR method has demonstrated the convergence
       speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-source
-    implementations of which are available;
+      implementations of which are available in web;
       \item the stochastic optimization methods inferior to the deterministic ones in the convergence
     speed and in reliability. It is manifested especially strongly on more complex multiextremal
-  problems;
-    \item the parallel version of the Globalizer system demonstrates good speedup
-  when running on several nodes, on each of which a single objective function value per iterations is
-  computed. When solving the problems with fast computable objective functions and using multiple
-  threads on the nodes, the speedup for the Globalizer system degrade with increasing
-  number of nodes;
-    \item the MIDACO system is the most suitable for simple global optimization problems with the
-  fast-computable objective functions. In this case, MIDACO is reliable enough and provides a linear
-  speedup with increasing number of nodes and threads executed on these ones in parallel.
+    problems.
     \end{itemize}
-    %Future work:
-    %\begin{itemize}
-    %  \item
-    %\end{itemize}
 \end{frame}
+
 {

### 2019-07-15 2019/MOTOR_Sovrasov/slides/slides.tex
MOTOR: add the final version of slides

@@
     \begin{itemize}
-      \item DIRECT, DIRECT$l$;
-      \item AGS, AGS$l$.
+      \item DIRECT, DIRECT$l$ (Gablonsky J. M. et al, 2001);
+      \item AGS (Strongin R.G., 1978).
     \end{itemize}
@@
     \begin{itemize}
-      \item Multi Level Single Linkage;
-      \item StoGO;
-      \item Differential Evolution;
-      \item Controlled Random Search;
-      \item Dual Simulated Annealing;
-      \item Ant Colony Optimization.
+      \item Multi Level Single Linkage (Alexander H. G. et al, 1987);
+      \item StoGO (Madsen, Kaj at al, 1998);
+      \item Differential Evolution (Storn, Rainer et al, 1997);
+      \item Controlled Random Search (Price, W. L., 1983);
+      \item Dual Simulated Annealing (Y Xiang et al, 1997);
     \end{itemize}
@@
   \frametitle{Univariate Algorithm of Global Search}
-  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
+  Optimization method generates search sequence \(\{x_k:x_k\in[a,b]\}\) and consists of the following steps:
   \begin{enumerate}
@@
     \item[Step 2.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
-    \item[Step 3.] Choose \(p\) intervals \((x_{t_j-1}, x_{t_j})\) with the greatest characteristics and
-    compute objective \(\varphi(y(x^{k+j}))\) in points chosen using the decision rule \(d\):
+    \item[Step 3.] Choose the interval \((x_{t-1}, x_{t})\) with the greatest characteristic and
+    compute objective \(\varphi(y(x^{k+1}))\) in the point chosen using the decision rule \(d\):
     \begin{displaymath}
-      x^{k+1+j}=d(t)\in (x_{t_j-1}, x_{t_j}),\:j=\overline{1,p}
+      x^{k+1}=d(t)\in (x_{t-1}, x_{t})
     \end{displaymath}
-    \item[Step 4.] If \(x_{t_j}-x_{t_j-1}<\varepsilon\) for one of \(j=\overline{1,p}\), stop the method.
+    \item[Step 4.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
   \end{enumerate}
@@
   \end{displaymath}
-  where $L$ is current estimation on Holder constant, $r$ is reliability coefficient.
+  where $L$ is the current estimation of the Holder constant, $r$ is reliability coefficient.
   AGS is very sensitive to choice of $r$:
@@
     \begin{column}{0.5\textwidth}
-      Generator GKLS was employed to construct 8 sets of 100 test problems:
+      Generator GKLS (Gaviano, M. et al, 2003) was employed to construct 8 sets of 100 test problems:
       \begin{displaymath}
@@
   \begin{itemize}
-    \item in the DIRECT and DIRECT\(l\) methods, the parameter \(\epsilon=10^{-4}\);
+    \item in the DIRECT and DIRECT\(l\) methods, parameter \(\epsilon=10^{-4}\);
     \item in the SDA method, the parameter \(visit=2.72\).
@@
     \caption{Operating characteristics of AGS and AGS-AR with different hyperparameters
-    when solving problems from the GKLS 5d Simple classe.}
+    when solving problems from the GKLS 5d Simple class.}
     \label{fig:stability}
@@
       performs on-par with properly tuned AGS;
-      \item AGS-AR method has demonstrated the convergence
-      speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-source
+      \item AGS-AR method has demonstrated convergence
+      speed and reliability at the level of DIRECT and exceeds many other algorithms, open-source
       implementations of which are available in web;
-      \item the stochastic optimization methods inferior to the deterministic ones in the convergence
+      \item considered stochastic optimization methods inferior to the deterministic ones in the convergence
     speed and in reliability. It is manifested especially strongly on more complex multiextremal

### 2019-08-05 2019/RSD_Integer/RSD2019.tex

@@
 
-\title{Parallel Global Optimization for Non-Convex Mixed-Integer Problems\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}}
+\title{Parallel Global Optimization for Non-Convex Mixed-Integer Problems
+%\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
+}
 \author{Konstantin Barkalov \and Ilya Lebedev %\Letter 
@@
 
-\keywords global optimization, non-convex constraints, mixed-integer problems.
+\keywords { Global optimization, Non-convex constraints, Mixed-integer problems}
 
@@
 
+\medskip
+
+\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,16-11-10150.
+
 \begin{thebibliography}{10}

### 2019-08-20 2019/Automation/Barkalov.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Sample document for preparing papers to  "Avtomatika i Telemekhanika"
%%  charset=windows-1251
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt]{a&t}
\usepackage{graphicx}

\begin{document}  %%%!!!

\year{2019}
\title{НАЗВАНИЕ СТАТЬИ ИЛИ ЗАМЕТКИ БЫВАЕТ С ФОРМУЛАМИ $a+b=c$}%
\thanks{Исследование выполнено за счет гранта Российского научного фонда
(проект \mbox{№\,16-11-10150}).}

\authors{Р.Г.~СТРОНГИН, д-р~физ.-мат.~наук\\
В.П.~ГГЕРГЕЛЬ, д-р~техн.~наук\\
К.А.~БАРКАЛОВ, канд.~физ.-мат.~наук\\
(Нижегородский государственный университет им. Н.И. Лобачевского)}

\maketitle

\begin{abstract}
Краткая аннотация статьи или заметки. Иногда не бывает. Хх
хххххххххххх хххххх хххххххх х хххххххххххххх ххх ххххххх хххх.
Хххххххххххххх ххх ххххх хххххх х хххххххххххххх ххх ххххххх
хххх.
\end{abstract}

\section{Введение}

Постановка задачи
Условие Липшица
Обзор методов Липшицевой оптимизации - взять из диссертации.

Информационно-статистический алгоритм - основы заложены Неймарком, развиты Стронгиным, обобщены учениками Стронгина.

В статье - последние наработки в данном направлении.
Редукция размерности - взять из WCGO

\section{Базовый алгоритм глобального поиска}

\section{Многошаговая схема редукции размерности}

Многошаговая схема редукции размерности (схема вложненной оптимизации) основана на известном соотношении
\begin{equation}\label{nested}
\min_{y \in D}\varphi(y) = \min_{y_1\in\left[a_1,b_1\right]}\min_{y_2\in\left[a_2,b_2\right]}...\min_{y_N\in\left[a_N,b_N\right]}\varphi(y),
\end{equation}
которое позволяет свести решение исходной многомерной задачи (\ref{main_problem}) к решению семейства рекурсивно связанных одномерных позадач.

Для формального описания многошаговой схемы введем семейство функций, определяемых в соответствии с соотношениями 
\begin{equation}\label{nested_N}
\varphi_N(y_1,...,y_N) \equiv \varphi(y_1,...,y_N),
\end{equation}
\begin{equation}\label{nested_i}
\varphi_i(y_1,...,y_i) = \min_{ y_{i+1} \in\left[a_{i+1},b_{i+1}\right]} \varphi_{i+1}(y_1,...,y_i,y_{i+1}), 1\leq i\leq N-1.
\end{equation}

Тогда, в соответствии с (\ref{nested}), для решения многомерной задачи (\ref{main_problem}) достаточно решить одномерную задачу  
\begin{equation}\label{nested_1}
\varphi^* = \min_{y_1\in\left[a_1,b_1\right]}\varphi_1(y_1).
\end{equation}
Однако каждое вычисление значения функции $\varphi_1$ в некоторой фиксированной точке $y_1$ предполагает решение одномерной задачи оптимизации второго уровня 
\begin{equation}
\varphi_1(y_1) = \min_{y_2\in\left[a_2,b_2\right]}\varphi_2(y_1,y_2).
\end{equation}
Вычисление значений функции $\varphi_2$, в свою очередь, требует одномерной минимизации функции $\varphi_3$, и т.д. вплоть до
решения задачи
\begin{equation}
\varphi_{N-1}(y_1,...,y_{N-1}) = \min_{ y_{N} \in\left[a_{N},b_{N}\right]} \varphi_{N}(y_1,...,y_{N}),
\end{equation}
на последнем уровне рекурсии.

Схема вложенной оптмизации (\ref{nested_N})--(\ref{nested_1}) в предположении липшицевости одномерных функций $\overline{\varphi}_i(y_i)=\varphi_i(y_1,...,y_i)$ послужила основой для обобщения методов одномерной оптимизации [7,25,32,34,36 из JOGO] на случай многих переменных [3,12,25,26,36 из JOGO].

Дадим более подробное описание взаимосвязей между одномерными подзадачами в многошаговой схеме. Для удобства последующего изложения введем обозначение
\[
v_i = (y_1,...y_i), \; 1 \leq i\ \leq N.
\]
Тогда одномерные функции $\overline{\varphi}_i(y_i), \; 1 < i\ \leq N$, порожденные в рамках многошаговой схемы, могут быть представлены в виде
\[
\overline{\varphi_i}(y)_i = \varphi_i(v_{i-1},y_i), \; 1 < i\ \leq N,
\]
где вектор $v_{i-1}$ фиксирован.

В соответствии с многошаговой схемой на некотором шаге поиска $i, 1 \leq i \leq k_1$  алгоритм начинает минимизировать одномерную функцию $\varphi_1(y_1)$. Для этого, используя накопленную поисковую информацию, алгоритм выбирает точку очередной итерации $\widehat{y}_1$ и вычисляет значение функции в ней. Но для того, чтобы получить значение $\varphi_1(\widehat{y}_1)$ в этой точке, должна быть решена задача минимизации функции $\varphi_2(\widehat{y}_1,y_2)$. То есть, алгоритм приостанавливает минимизацию функции $\varphi_1(y_1)$ и начинает минимизировать $\varphi_2(\widehat{y}_1,y_2)$. В процессе минимизации алгоритм генерирует последовательность точек $\{y_2^j, 1 \leq j \leq k_2\}$, для каждой из которых необходимо решить задачу минимизации функции $\varphi_3(\widehat{y}_1,\widehat{y}_2,y_3)$, и т.д. Данный процесс выполняется рекурсивно до последнего $N$-го уровня. Возникающие при этом подзадачи и взаимосвязи между ними отображены на рис. \ref{fig1}.

\begin{figure}
\begin{center}%[scale=0.3][width=0.75\textwidth]
  \includegraphics[width=0.9\textwidth]{fig1.png} 
  \caption{Взаимосвязи между подзадачами в многошаговой схеме}\label{fig1} 
\end{center}
\end{figure}

Наглядно видно, что структура взаимосвязей имеет форму дерева. Указанная структура может меняться в процессе решения многомерной задачи (\ref{main_problem}): вычисление значения функции $\varphi_i(\widehat{v}_{i-1},y_i), \; 1\leq i < N,$  на $i$-м уровне рекурсии требует решения всех подзадач в одном из поддеревьев $(i+1)$-го уровня. При этом функции $\varphi_N(\widehat{v}_{N-1},y_N)$ на $N$-м уровне являются листьями дерева задач, их значения вычисляются непосредственно.

%недостатки многошаговой схемы

\section{Адаптивная многошаговая схема}

Рассмотрим обобщение схемы вложенной оптимизации (\ref{nested_N})--(\ref{nested_1}), позволяющее преодолеть недостатки указанные выше. Основным отличием является отказ от принципа подчитенности одномерных подзадач на основе функций $\overline{\varphi_i}(y_i), \; 1\leq i \leq N,$ порожденных в рамках многошаговой схемы. В новом подходе все возникающие подзадачи предлагается решать одновременно. Предварительное описание \textit{адаптивной многошаговой схемы} редукции размерности заключается в следующем.

Введем следующее определение.

\begin{definition}
Хххххххххххххх хххххх-хххххххх х ххххх ххххххххх ххх ххххххх хххх.
Ххххх хххххххх хххххххххххххх х ххххх.
\end{definition}

Хххх ххххххх хххххххххххххх х ххххх ххххххххх ххх ххххххх хххх.

Рассмотрим следующую задачу.

\begin{problem} \label{prob:1}
Хххххх хххххххх хххххх хххххххх х ххххх ххххххххх ххх ххххххх хххх.
Ххххх хххххххх хххххххххххххх х ххххххххх х ххххх.
\end{problem}

Хххххх хххххххх хххххх хххххххх х ххххх ххххххххх ххх ххххххх хххх.
Ххххх хххххххх хххххххххххххх х ххххххххх х ххххх.

\section{Заголовок третьего раздела}

\subsection{Заголовок подраздела}

Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.

Сформулируем следующую теорему.

\begin{theorem}[{\cite[c.\,123]{first}}] %
Пусть выполнены следующие условия:

\begin{ruslist}
\item
первое условие;

\item
второе условие.
\end{ruslist}

Тогда справедливы следующие утверждения:

\begin{enumlist}
\item
первое утверждение;

\item
второе утверждение.
\end{enumlist}
\end{theorem}

\begin{proof}
Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх:
\begin{gather}
    2\times 2=4.
\end{gather}
Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
\end{proof}

Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.

\begin{corollary}
Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
Хххххххххххххх хххххххх х хххххххххххххх ххх ххххххх хххх.
\end{corollary}

Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.

\subsection{Заголовок следующего подраздела}

Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.

\begin{lemma}[(см.\ {\cite[c.\,45]{second}})] \label{lm:1}
Хххххххххххххх хххххххххххххх х хххххххххх:
\begin{multline}
    2\times 2\times 2\times 2\times 2\times 2\times 2\times 2\times
    2\times 2\times 2\times 2\times 2\times 2\times 2\times 2\times
\\
    \times 2\times 2\times 2\times 2\times 2\times 2\times 2\times 2\times
    2\times 2\times 2\times 2\times 2\times 2\times 2\times 2
\end{multline}
хххххххх ххххххх ххххххх х хххххххххххххх ххх.
\end{lemma}

Опишем следующий алгоритм.

\begin{algorithm}[(Быстрый)] \label{alg:1}
\ %%<-- этот пробел для того, чтобы первый элемент перечня был
%% на следующей строке, а не в подбор к заголовку окружения

\begin{enumlist}[.] % перечни, нумеруемые 1. 2. и т.д.
% \setcounter{enumlisti}{-1} % <-- эта команда нужна
%% для нумерации элементов перечня с нулевого
\item
Начать.

\item
Изменить.

\item
Закончить.
\end{enumlist}
\end{algorithm}

Следующая теорема утверждает сходимость алгоритма~\ref{alg:1}.

\begin{theorem}[(Теорема сходимости)] \label{th:2}
Алгоритм~{\rm\ref{alg:1}} сходится.
\end{theorem}

Можно привести замечание.

\begin{remark}
Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
\end{remark}

\section{Четвертый раздел}

Данный раздел содержит несколько примеров различных окружений.

\begin{example}
Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.

Некоторые перечни можно нумеровать русскими буквами:
\begin{ruslist} % перечни, нумеруемые а), б) и т.д.
\item
ххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хх
ххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх;

\item
ххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
\end{ruslist}

А некоторые можно~"--- и латинскими буквами:

\begin{latlist} % перечни, нумеруемые а), б) и т.д.
\item
ххххххххххх ххх ххххххх хххх;

\item
хххххххххххх ххх ххххххх хххх.
\end{latlist}
Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
\end{example}

Можно сформулировать утвеждение.

\begin{statement}
Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
\end{statement}

Можно сформулировать предложение.

\begin{proposition}
Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
\end{proposition}

\appendix{1}  % приложения можно нумеровать, если их несколько

\begin{proofoftheorem}{\ref{th:2}}
Докажем сначала, что выполнено следующее соотношение:
\begin{gather} \label{f(x)=y}
    f(x)=y.
\end{gather}
Действительно,~\dots откуда получаем, что
равенство~\eqref{f(x)=y} справедливо. Следовательно,~\dots
окончательно,~\dots

Теорема~\ref{th:2} доказана.
\end{proofoftheorem}

\begin{proofoflemma}{\ref{lm:1}}
Очевидно, что \dots Таким образом, \dots, что и требовалось
доказать.
\end{proofoflemma}

\begin{thebibliography}{10}

\bibitem{Neymark66a}
{\it Неймарк Ю.И., Стронгин Р.Г.} Поиск экстремума функций по принципу максимума информации // АиТ. 1966. № 11. С. 113--118.

\bibitem{Neymark66b}
{\it Неймарк Ю.И., Стронгин Р.Г.} Информационный подход к задаче поиска экстремума функций // Изв. АН СССР, Техническая кибернетика. 1966. № 1. С. 17--26.

\bibitem{Strongin2000}
{\it Strongin R.G., Sergeev Ya.D.}
Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Dordrecht: Kluwer Academic Publishers, 2000.

\bibitem{Strongin2013}
{\it Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А.}
Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013.

\bibitem{second}
{\it Второй В.В., Первый П.П.}
Монография по данной теме. М.: Высш. шк., 1999.

\bibitem{third}
{\it Третий Т.Т.}
Публикация в трудах конференции //
Тр. Ин-та такого-то РАН. 2000. Т. 1. № 2. С. 3--4.

\bibitem{forth}
{\it Четвертый Ч.Ч.}
Публикация по теме в серийном издании или сборнике /
Сб. научн. тр. к 20-летию Института. Новосибирск, Изд-во <<Пирогов>>, 2000. Т. 1. № 2. С. 3--4.

\bibitem{fifth}
{\it Fifth F.}
Some journal publication in English //
Appl. Math. Comput. J., Elsevier Publ. 1925. V. 501. No. 1. P. 1234--5678.

\bibitem{sixth}
{\it Sixth J.Th.}
A book in English. Boston: Springer, 1991.

\end{thebibliography}

\AdditionalInformation{Стронгин Р.Г.}{Нижегородский государственный университет им. Н.И. Лобачевского, президент, Нижний Новгород}{strongin@unn.ru}

\AdditionalInformation{Гергель В.П.}{Нижегородский государственный университет им. Н.И. Лобачевского, директор института информационных технологий, математики и механики, Нижний Новгород}{gergel@unn.ru}

\AdditionalInformation{Баркалов К.А.}{Нижегородский государственный университет им. Н.И. Лобачевского, доцент кафедры математического обеспечения и суперкомпьютерных технологий института информационных технологий, математики и механики, Нижний Новгород}{konstantin.barkalov@itmm.unn.ru}

\end{document}

### 2019-08-20 2019/Automation/Barkalov.tex
Introduction

@@
 \title{НАЗВАНИЕ СТАТЬИ ИЛИ ЗАМЕТКИ БЫВАЕТ С ФОРМУЛАМИ $a+b=c$}%
-\thanks{Исследование выполнено за счет гранта Российского научного фонда
-(проект \mbox{№\,16-11-10150}).}
+\thanks{Исследование выполнено за счет гранта Российского научного фонда (проект \mbox{№\,16-11-10150}).}
 
@@
 
-Постановка задачи
-Условие Липшица
+В статье рассматриваются задачи глобальной оптимизации вида
+\begin{eqnarray}\label{main_problem}
+& \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
+& D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
+\end{eqnarray}
+Предполагается, что целевая функция может быть многоэкстремальной, задана неявно (функция вида ``черный ящик''), а вычисление ее значений связано с решением задачи численного моделирования и является трудоемкой операцией.
+
+Любая возможность достоверно оценить глобальный оптимум в многоэкстремальной задаче с функциями вида ``черный ящик'' принципиально связана априорной информацией, позволяющей связать возможные значения целевой функции с ее известными значениями в точках проведенных испытаний.
+Для многих прикладных задачах типичной является ситуация, когда ограниченное изменение вектора параметров $y$ вызывает ограниченное изменение значений $\varphi(y)$.
+Математической моделью, описывающей указанное предположение, является условие Липшица
+\[
+\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in D,\; 0<L<\infty.
+\]
+Отметим, что значение константы Липшица $L$ обычно является априори неизвестным, что делает оценку константы одной из ключевых проблем при построении методов липшицевой оптимизации.
+
 Обзор методов Липшицевой оптимизации - взять из диссертации.
+
+Одним из эффективных детерминированных методов решения задач многоэкстремальной оптимизации является \textit{информационно-статистический алгоритм глобального поиска}.
+Основы информационно-статистического подхода были заложены Ю.И.
+Неймарком \cite{Neymark66a,Neymark66b} и развиты Р.Г.
+Стронгиным \cite{Strongin78}.
+Впоследствии метод, изначально предложенный для решения безусловных задач, был обобщен для решения задач с невыпуклыми ограничениями \cite{Strongin86,Strongin87} и многокритериальных задач \cite{Strongin93}.
 
@@
 
-\bibitem{Strongin2000}
-{\it Strongin R.G., Sergeev Ya.D.}
-Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Dordrecht: Kluwer Academic Publishers, 2000.
+\bibitem{Strongin78}
+{\it Стронгин Р.Г.} Численные методы в многоэкстремальных задачах (информационно-статистические алгоритмы). М.: Наука, 1978.
 
-\bibitem{Strongin2013}
-{\it Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А.}
-Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013.
+\bibitem{Strongin86}
+{\it Стронгин Р.Г., Маркин Д.Л.} Минимизация многоэкстремальных функций при невыпуклых ограничениях // Кибернетика. 1986. №4. С.63--69.
 
-\bibitem{second}
-{\it Второй В.В., Первый П.П.}
-Монография по данной теме. М.: Высш. шк., 1999.
+\bibitem{Strongin87}
+{\it Маркин Д.Л., Стронгин Р.Г.} Метод решения многоэкстремальных задач с невыпуклыми ограничениями, использующий априорную информацию об оценках оптимума // Ж.
+вычисл.
+матем.
+и матем.
+физ.
+1987.
+Т.27, № 1.
+С.52--61.
+
+\bibitem{Strongin93}
+{\it Маркин Д.Л., Стронгин Р.Г.} О равномерной оценке множества слабоэффективных точек в многоэкстремальных многокритериальных задачах оптимизации // Ж.
+вычисл.
+матем.
+и матем.
+физ.
+1993 Т.
+33, № 2.
+С.
+195--205.
+
+\bibitem{Strongin00}
+{\it Strongin R.G., Sergeev Ya.D.} Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Dordrecht: Kluwer Academic Publishers, 2000.
+
+\bibitem{Strongin13}
+{\it Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013.
 

### 2019-08-20 2019/Automation/Barkalov.tex
Correct introduction

@@
 \end{eqnarray}
-Предполагается, что целевая функция может быть многоэкстремальной, задана неявно (функция вида ``черный ящик''), а вычисление ее значений связано с решением задачи численного моделирования и является трудоемкой операцией.
+Предполагается, что целевая функция может быть многоэкстремальной, задана неявно (функция вида <<черный ящик>>), а вычисление ее значений связано с решением задачи численного моделирования и является трудоемкой операцией.
 
-Любая возможность достоверно оценить глобальный оптимум в многоэкстремальной задаче с функциями вида ``черный ящик'' принципиально связана априорной информацией, позволяющей связать возможные значения целевой функции с ее известными значениями в точках проведенных испытаний.
+Любая возможность достоверно оценить глобальный оптимум в многоэкстремальной задаче с функциями вида <<черный ящик>> принципиально основана на априорной информации, позволяющей связать возможные значения целевой функции с ее известными значениями в точках проведенных испытаний.
 Для многих прикладных задачах типичной является ситуация, когда ограниченное изменение вектора параметров $y$ вызывает ограниченное изменение значений $\varphi(y)$.
@@
 \]
-Отметим, что значение константы Липшица $L$ обычно является априори неизвестным, что делает оценку константы одной из ключевых проблем при построении методов липшицевой оптимизации.
+%Отметим, что значение константы $L$ обычно является априори неизвестным, что делает ее оценку одной из ключевых проблем при построении методов липшицевой оптимизации.
+Предположение липшицевости целевой функции типично для многих подходов к разработке оптимизационных алгоритмов.
+Первые методы липшицевой оптимизации были предложены в начале 70-х годов \cite{Evtushenko71,Pijavski72,Shubert72}; с тех пор данное направление продолжает активно развиваться \cite{Jones93,Pinter96,Zilinskas08,Evtushenko07,Evtushenko09}.
+Например, многие известные методы основаны на различных способах разбиения области поиска на систему подобластей и последующего выбора наиболее перспективной подобласти для размещения очередного испытания (вычисления целевой функции).
+Результаты в данном направлении представлены в работах \cite{Jones09,Zilinskas10,Evtushenko13,Kvasov13,Paulavicius16}.
 
-Обзор методов Липшицевой оптимизации - взять из диссертации.
+В настоящее время для решения задач оптимизации с функциями вида <<черный ящик>> широко используются генетические и популяционные алгоритмы (см., например, \cite{Kureichik04,Karpenko08}), которые так или иначе основаны на идеях случайного поиска.
+В силу простоты реализации и использования они получили большую популярность, однако по качеству работы (численной оценкой которого может служить число корректно решенных задач из некоторого набора) они существенно уступают детерминированным алгоритмам \cite{Kvasov18,Sergeyev18}.
+
+%В последние годы для решения задач с функциями вида ``черный ящик'' получили широкое распространение так называемые биоинспирированные алгоритмы, основанные на моделировании природных процессов и явлений.
+Типичными примерами методов данного класса являются генетические и популяционные алгоритмы (см., например, ).
+В основе указанных алгоритмов так или иначе лежит случайный поиск, т.к.
+они все содержат случайную компоненту (например, операции скрещивания и мутации в генетических или миграции в популяционных алгоритмах).
+В силу своей вероятностной природы методы данного класса обладают, вообще говоря, всюду плотной сходимостью, что невыгодно отличает их от детерминированных алгоритмов.
 
@@
 Стронгиным \cite{Strongin78}.
-Впоследствии метод, изначально предложенный для решения безусловных задач, был обобщен для решения задач с невыпуклыми ограничениями \cite{Strongin86,Strongin87} и многокритериальных задач \cite{Strongin93}.
+Впоследствии метод, изначально предложенный для решения безусловных задач, был успешно обобщен для решения задач с невыпуклыми ограничениями \cite{Strongin86,Strongin87} и многокритериальных задач \cite{Strongin93}.
 
-Информационно-статистический алгоритм - основы заложены Неймарком, развиты Стронгиным, обобщены учениками Стронгина.
+Разработанные методы основаны на редукции исходной многомерной задачи к эквивалентной одномерной или системе одномерных подзадач и последующим решением одномерных задач эффективными методами оптимизации функций одной переменной.
+Предложено две такие схемы: редукция на основе кривых, заполняющих пространство (кривых Пеано, или разверток) \cite{Strongin91,Strongin00}, и схема рекурсивной вложенной оптимизации (многошаговая схема) \cite{Grishagin97,Grishagin01}.
+В статье \cite{Grishagin16} предложена адаптивная многошаговая схема, существенно повышающая эффективность оптимизации по сравнению с базовым прототипом.
+В данной работе предлагается обобщение адаптивной схемы редукции размерности, комбинирующее использование вложенной оптимизации и кривых Пеано.
+При таком подходе вложенные подзадачи в адаптивной схеме могут быть как одномерными, так и многомерными.
+В последнем случае для редукции размерности вложенных подзадач используются развертки.
 
-В статье - последние наработки в данном направлении.
-Редукция размерности - взять из WCGO
+%Развитие - параллельность.
+предложены параллельные варианты многих алгоритмов.
+Отметим, что в основном распараллеливают генетические алгоритмы.
+В статье - последние наработки в направлении алгоритма глобального поиска.
 
@@
 
+\bibitem{Evtushenko71}
+{\it	Евтушенко Ю.Г.} Численный метод поиска глобального экстремума функций (перебор на неравномерной сетке) // Ж. вычисл. матем. и матем. физ. 1971. T. 11, №6, С. 1390--1403.
+
+\bibitem{Pijavski72}
+{\it	Пиявский С.А.} Один алгоритм отыскания абсолютного экстремума функций // Ж. вычисл. матем. и матем. физ. 1972. Т. 12, № 4. С. 888--896.
+
+\bibitem{Shubert72}
+{\it Shubert B.O.} A sequential method seeking the global maximum of a function // SIAM J. Numer. Anal. 1972. V. 9. P. 379–388.
+
+\bibitem{Jones93} 
+{\it Jones D.R., Perttunen C.D., Stuckman B.E.} Lipschitzian optimization without the Lipschitz constant // J. Optim. Theory Appl. 1993. V. 79. No. 1. P. 157--181.
+
+\bibitem{Pinter96}
+{\it Pinter J. D.} Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications). Dordrecht: Kluwer Academic Publishers, 1996.
+
+\bibitem{Zilinskas08}
+{\it {\v Z}ilinskas J.} Branch and bound with simplicial partitions for global optimization // Math. Model. Anal. 2008. V. 13. No. 1. P. 145--159.
+
+\bibitem{Evtushenko07}
+{\it Евтушенко Ю.Г., Малкова В.У., Станевичюс А.-И. А.} Распараллеливание процесса поиска глобального экстремума // Автомат. и телемех. 2007. № 5. C. 46--58
+
+\bibitem{Evtushenko09}
+{\it Евтушенко Ю.Г., Малкова В.У., Станевичюс А.-И. А.} Параллельный поиск глобального экстремума функций многих переменных // Ж. вычисл. матем. и матем. физ. 2009. Т. 49, №2. C. 255--269.
+
+\bibitem{Jones09}%Проверить формат ссылки!!!
+{\it Jones D. R.} The direct global optimization algorithm. In: Floudas C. A., Pardalos P. M. (eds.) The Encyclopedia of Optimization, Second Edition.  Springer. 2009. P. 725--735.
+
+\bibitem{Zilinskas10}
+{\it Paulavi{\v c}ius R., {\v Z}ilinskas J., Grothey A.}: Investigation of selection strategies in branch and bound algorithm with simplicial partitions and combination of Lipschitz bounds // Optim.
+Lett.
+2010.
+V.
+4(1).
+P.
+173–--83
+
+\bibitem{Evtushenko13}
+{\it Evtushenko Y. G., Posypkin M. A.} A deterministic approach to global box-constrained optimization // Optim. Lett. 2013. V. 7(4). P. 819--829
+
+\bibitem{Kvasov13}
+{\it Квасов Д.Е., Сергеев Я.Д.} Методы липшицевой глобальной оптимизации в задачах управления // Автомат. и телемех. 2013. № 9. C. 3--19.
+
+\bibitem{Paulavicius16}
+{\it Paulavi{\v c}ius R., {\v Z}ilinskas J.} Advantages of simplicial partitioning for Lipschitz optimization problems with linear constraints // Optim. Lett. 2016. V. 10(2). P. 237--246.
+
+\bibitem{Kureichik04}
+{\it Гладков Л.А., Курейчик В.В., Курейчик В.М.} Генетические алгоритмы. Учебное пособие. М.: ФИЗМАТЛИТ, 2004.
+
+\bibitem{Karpenko08}
+{\it Карпенко А.П.} Современные алгоритмы поисковой оптимизации. Алгоритмы, вдохновленные природой: учебное пособие. М.: Издательство МГТУ им. Н. Э. Баумана, 2014.
+
+\bibitem{Kvasov18}
+{\it Kvasov D.E., Mukhametzhanov M.S.} Metaheuristic vs. deterministic global optimization algorithms: The univariate case // Appl. Math. Comput. 2018. V. 318. P. 245--259.
+
+\bibitem{Sergeyev18}
+{\it Sergeyev Y., Kvasov D., Mukhametzhanov M.} On the efficiency of nature-inspired metaheuristics in expensive global optimization with limited budget // Sci.
+Rep. V. 8(1). Art. No. 435.
+
 \bibitem{Neymark66a}
-{\it Неймарк Ю.И., Стронгин Р.Г.} Поиск экстремума функций по принципу максимума информации // АиТ. 1966. № 11. С. 113--118.
+{\it Неймарк Ю.И., Стронгин Р.Г.} Поиск экстремума функций по принципу максимума информации // Автомат. и телемех. 1966. № 11. С. 113--118.
 
@@
 
+\bibitem{Strongin91}
+{\it Стронгин Р.Г.} Параллельная многоэкстремальная оптимизация с использованием множества разверток // Ж. вычисл. матем. и матем. физ. 1991. T. 31, № 8. С. 1173--1185.
+
 \bibitem{Strongin00}
@@
 
-\bibitem{third}
-{\it Третий Т.Т.}
-Публикация в трудах конференции //
-Тр. Ин-та такого-то РАН. 2000. Т. 1. № 2. С. 3--4.
+\bibitem{Grishagin97}
 
-\bibitem{forth}
-{\it Четвертый Ч.Ч.}
-Публикация по теме в серийном издании или сборнике /
-Сб. научн. тр. к 20-летию Института. Новосибирск, Изд-во <<Пирогов>>, 2000. Т. 1. № 2. С. 3--4.
+{\it Grishagin V.A., Sergeyev Y.D., Strongin R.G.} Parallel characteristical algorithms for solving problems of global optimization // J. Glob. Optim. 1997. V. 10(2). P. 185--206.
 
-\bibitem{fifth}
-{\it Fifth F.}
-Some journal publication in English //
-Appl. Math. Comput. J., Elsevier Publ. 1925. V. 501. No. 1. P. 1234--5678.
+\bibitem{Grishagin01}
+{\it Sergeyev Y., Grishagin V.} Parallel asynchronous global search and the nested optimization scheme // J. Comput. Anal. Appl. 2001. V. 3(2). P. 123--145.
 
-\bibitem{sixth}
-{\it Sixth J.Th.}
-A book in English. Boston: Springer, 1991.
+\bibitem{Grishagin16}
+{\it Gergel V., Grishagin V., Gergel A.} Adaptive nested optimization scheme for multidimensional global search // J. Glob. Optim. 2016. V. 66(1). P. 35--51.
+
+%\bibitem{Sergeyev00} Kvasov, D.E., Pizzuti, C., Sergeyev, Y.D.: Local tuning and partition strategies for diagonal GO methods. Numer. Math. 94(1), 93--106 (2003)
+%\bibitem{Sergeyev06} Sergeyev, Ya.D., Kvasov, D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim. 16(3), 910--937 (2006)
+
+%\bibitem{third} {\it Третий Т.Т.} Публикация в трудах конференции // Тр. Ин-та такого-то РАН. 2000. Т. 1. № 2. С. 3--4.
+%\bibitem{forth} {\it Четвертый Ч.Ч.} Публикация по теме в серийном издании или сборнике / Сб. научн. тр. к 20-летию Института. Новосибирск, Изд-во <<Пирогов>>, 2000. Т. 1. № 2. С. 3--4.
+%\bibitem{fifth} {\it Fifth F.} Some journal publication in English // Appl. Math. Comput. J., Elsevier Publ. 1925. V. 501. No. 1. P. 1234--5678.
+%\bibitem{sixth} {\it Sixth J.Th.} A book in English. Boston: Springer, 1991.
 

### 2019-08-21 2019/Automation/Barkalov.tex
Numerical experiments

@@
 Возникающие при этом подзадачи и взаимосвязи между ними отображены на рис.
-\ref{fig1}.
+\ref{fig0}.
 
@@
   \includegraphics[width=0.9\textwidth]{fig1.png} 
-  \caption{Взаимосвязи между подзадачами в многошаговой схеме}\label{fig1} 
+  \caption{Взаимосвязи между подзадачами в многошаговой схеме}\label{fig0} 
 \end{center}
@@
 
-Введем следующее определение.
+%взять первый вариант из статьи в JoGO
+\section{Результаты экспериментов}
 
-\begin{definition}
-Хххххххххххххх хххххх-хххххххх х ххххх ххххххххх ххх ххххххх хххх.
-Ххххх хххххххх хххххххххххххх х ххххх.
-\end{definition}
+Стандартный подход к сравнению алгоритмов глобальной оптимизации основан на решении указанными методами серии задач, выбранных случайно из некоторого класса.
+Будем использовать два класса многоэкстремальных тестовых функций (GKLS \textit{Simple}, GKLS \textit{Hard}), описанные в \cite{Gaviano03}.
 
-Хххх ххххххх хххххххххххххх х ххххх ххххххххх ххх ххххххх хххх.
+В работах \cite{Grishagin18,Sovrasov19} было показано, что алгоритм глобального поиска (АГП) как с использованием разверток, так и в сочетании с адаптивной схемой редукции размерности превосходит многие известные алгоритмы глобальной оптимизации, включая методы DIRECT и DIRECT\textit{l}.
+Поэтому в данном исследовании мы ограничимся сравнением вариантов АГП в сочетании с различными схемами редукции размерности.
 
-Рассмотрим следующую задачу.
+Для сравнения эффективности работы алгоритмов будем использовать два критерия: среднее число испытаний и операционные характеристики.
+Операционной характеристикой алгоритма называется функция $P(k)$, определяемая как доля задач из рассматриваемой серии, для решения которых потребовалось не более чем $k$ испытаний.
+Задачу будем считать решенной, если алгоритм сгенерировал точку $y^k$ очередного испытания в окрестности решения $y^\ast$, т.е.
+$\left|y^k-y^\ast\right| < \delta \left\|b-a\right\|$, где $\delta = 0.01$, $a$ и $b$ --- границы области поиска $D$.
 
-\begin{problem} \label{prob:1}
-Хххххх хххххххх хххххх хххххххх х ххххх ххххххххх ххх ххххххх хххх.
-Ххххх хххххххх хххххххххххххх х ххххххххх х ххххх.
-\end{problem}
+Первая серия экспериментов была проведена на двумерных задачах классов GKLS \textit{Simple} и GKLS \textit{Hard} (100 функций каждого класса).
+В табл.~\ref{tab1}  представлено среднее число испытаний, выполненных АГП с использованием разверток ($K_e$), многошаговой схемы ($K_n$) и адаптивной многошаговой схемы ($K_a$).
+На рис.~\ref{fig2}(a,b) приведены операционные характеристики алгоритмов, полученные на указанных классах задач.
+Непрерывная линия соответствует алгоритму с использованием разверток, короткий пунктир --- адаптивной многошаговой схемы, длинный пунктир -- многошаговой схемы.
+Результаты экспериментов показывают, что АГП с использованием адаптивной многошаговой схемы показывает примерно одинаковое быстродействие  по сравнению с развертками, и оба они значительно превосходят алгоритм, использующий многошаговую схему.
+Поэтому в дальнейших экспериментах мы ограничимся сравнением различных вариантов адаптивной схемы редукции размерности.
 
-Хххххх хххххххх хххххх хххххххх х ххххх ххххххххх ххх ххххххх хххх.
-Ххххх хххххххх хххххххххххххх х ххххххххх х ххххх.
+\begin{table}
+\centering
+\caption{Среднее число испытаний для двумерных задач}\label{tab1}
+\begin{tabular}{lccc}
+\hline\noalign{\smallskip}
+ & GKLS \textit{Simple} &  GKLS \textit{Hard} \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+ $K_e$ &  252 & 674 \\
+ $K_n$ &  697 & 1252 \\
+ $K_a$ &  279 & 815 \\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
 
-\section{Заголовок третьего раздела}
+\begin{figure}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{2DSimple.pdf} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{2DHard.pdf} \\ (b)}
+\end{minipage}
+\caption{Operating characteristics using 2d GKLS \textit{Simple} (a) and \textit{Hard} (b) classes}
+\label{fig2}
+\end{figure}
 
-\subsection{Заголовок подраздела}
+Вторая серия экспериментов была проведена на четырехмерных задачах классов GKLS \textit{Simple} и GKLS \textit{Hard} (100 функций каждого класса).
+В табл.~\ref{tab2}  представлено среднее число испытаний, выполненных АГП с использованием адаптивной многошаговой схемы ($K_a$) и блочной адаптивной схемы ($K_ba$) с формированием двух уровней подзадач одинаковой размерности $N_1=N_2=2$.
+Отметим, что при использовании исходного варианта адаптивной многошаговой схемы при решении задачи размерности $N=4$ формируется четыре уровня одномерных подзадач, что усложняет их обработку.
 
-Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
+На рас.~\ref{fig3}(a,b) приведены операционные характеристики алгоритмов, полученные на классах GKLS \textit{Simple} и GKLS \textit{Hard}.
+Пунктирная линия соответствует алгоритму с использованием адаптивной, а непрерывная ---  блочной адаптивной схемы.
+Результаты экспериментов показывают, что использование блочной адаптивной схемы дает ощутимый выигрыш по числу испытаний (до 35\%) по сравнению с исходной схемой редукции размерности.
 
-Сформулируем следующую теорему.
+\begin{table}
+\centering
+\caption{Среднее число испытаний для четырехмерных задач}\label{tab2}
+\begin{tabular}{lcc}
+\hline\noalign{\smallskip}
+ &    GKLS \textit{Simple} &  GKLS \textit{Hard} \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+ $K_a$  &  21 747 & 35 633 \\
+ $K_{ba}$ &  15 942 & 33 206 \\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
 
-\begin{theorem}[{\cite[c.\,123]{first}}] %
-Пусть выполнены следующие условия:
+\begin{figure}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{4DSimple.pdf} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{4DHard.pdf} \\ (b)}
+\end{minipage}
+\caption{Операционные характеристики на классах задач GKLS \textit{Simple} (a) и \textit{Hard} (b)}
+\label{fig3}
+\end{figure}
 
-\begin{ruslist}
-\item
-первое условие;
+\section{Заключение}
 
-\item
-второе условие.
-\end{ruslist}
-
-Тогда справедливы следующие утверждения:
-
-\begin{enumlist}
-\item
-первое утверждение;
-
-\item
-второе утверждение.
-\end{enumlist}
-\end{theorem}
-
-\begin{proof}
-Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх:
-\begin{gather}
-    2\times 2=4.
-\end{gather}
-Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-\end{proof}
-
-Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-
-\begin{corollary}
-Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-Хххххххххххххх хххххххх х хххххххххххххх ххх ххххххх хххх.
-\end{corollary}
-
-Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-
-\subsection{Заголовок следующего подраздела}
-
-Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-
-\begin{lemma}[(см.\ {\cite[c.\,45]{second}})] \label{lm:1}
-Хххххххххххххх хххххххххххххх х хххххххххх:
-\begin{multline}
-    2\times 2\times 2\times 2\times 2\times 2\times 2\times 2\times
-    2\times 2\times 2\times 2\times 2\times 2\times 2\times 2\times
-\\
-    \times 2\times 2\times 2\times 2\times 2\times 2\times 2\times 2\times
-    2\times 2\times 2\times 2\times 2\times 2\times 2\times 2
-\end{multline}
-хххххххх ххххххх ххххххх х хххххххххххххх ххх.
-\end{lemma}
-
-Опишем следующий алгоритм.
-
-\begin{algorithm}[(Быстрый)] \label{alg:1}
-\ %%<-- этот пробел для того, чтобы первый элемент перечня был
-%% на следующей строке, а не в подбор к заголовку окружения
-
-\begin{enumlist}[.] % перечни, нумеруемые 1. 2. и т.д.
-% \setcounter{enumlisti}{-1} % <-- эта команда нужна
-%% для нумерации элементов перечня с нулевого
-\item
-Начать.
-
-\item
-Изменить.
-
-\item
-Закончить.
-\end{enumlist}
-\end{algorithm}
-
-Следующая теорема утверждает сходимость алгоритма~\ref{alg:1}.
-
-\begin{theorem}[(Теорема сходимости)] \label{th:2}
-Алгоритм~{\rm\ref{alg:1}} сходится.
-\end{theorem}
-
-Можно привести замечание.
-
-\begin{remark}
-Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-\end{remark}
-
-\section{Четвертый раздел}
-
-Данный раздел содержит несколько примеров различных окружений.
-
-\begin{example}
-Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-Хххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-
-Некоторые перечни можно нумеровать русскими буквами:
-\begin{ruslist} % перечни, нумеруемые а), б) и т.д.
-\item
-ххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хх
-ххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх;
-
-\item
-ххххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-\end{ruslist}
-
-А некоторые можно~"--- и латинскими буквами:
-
-\begin{latlist} % перечни, нумеруемые а), б) и т.д.
-\item
-ххххххххххх ххх ххххххх хххх;
-
-\item
-хххххххххххх ххх ххххххх хххх.
-\end{latlist}
-Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-\end{example}
-
-Можно сформулировать утвеждение.
-
-\begin{statement}
-Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-\end{statement}
-
-Можно сформулировать предложение.
-
-\begin{proposition}
-Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-Ххххххххххх хххххххххххххх х хххххххххххххх ххх ххххххх хххх.
-\end{proposition}
-
-\appendix{1}  % приложения можно нумеровать, если их несколько
-
-\begin{proofoftheorem}{\ref{th:2}}
-Докажем сначала, что выполнено следующее соотношение:
-\begin{gather} \label{f(x)=y}
-    f(x)=y.
-\end{gather}
-Действительно,~\dots откуда получаем, что
-равенство~\eqref{f(x)=y} справедливо. Следовательно,~\dots
-окончательно,~\dots
-
-Теорема~\ref{th:2} доказана.
-\end{proofoftheorem}
-
-\begin{proofoflemma}{\ref{lm:1}}
-Очевидно, что \dots Таким образом, \dots, что и требовалось
-доказать.
-\end{proofoflemma}
+В данной работе предложена обобщенная адаптивная схема редукции размерности для задач глобальной оптимизации, комбинирующая использование кривых Пеано и схему вложенной (рекурсивной) оптимизации.
+Для решения редуцированных подзадач меньшей размерности используется алгоритм глобального поиска.
+Приведена вычислительная схема алгоритма, рассмотрены основные вопросы, связанные с использованием адаптивной схемы редукции размерности.
+Проведены вычислительные эксперименты на серии тестовых задач с целью сравнения эффективности различных схем редукции размерности.
+Результаты экспериментов показывают, что использование блочной адаптивной схемы редукции размерности может значительно сократить число испытаний, необходимое для решения задачи с заданной точностью.
 
@@
 
+\bibitem{Gaviano03}
+{\it Gaviano M., Lera D., Kvasov D.E., Sergeyev Ya.D.} Software for generation of classes of test functions with known local and global minima for global optimization // ACM Trans.
+Math.
+Software.
+2003.
+V.
+29.
+P.
+469--480.
+
+\bibitem{Grishagin18}
+{\it Grishagin V., Israfiov R., Sergeyev Y.} Convergence conditions and numerical comparison of global optimization methods based on dimensionality reduction schemes // Appl.
+Math.
+Comput.
+2018.
+V.
+318.
+P.
+270--280.
+
+\bibitem{Sovrasov19}
+{\it Sovrasov V.} Comparison of several stochastic and deterministic derivative-free global optimization algorithms // Lecture Notes in Computer Science. 2019. V. 11548. P. 70--81.
+
 %\bibitem{Sergeyev00} Kvasov, D.E., Pizzuti, C., Sergeyev, Y.D.: Local tuning and partition strategies for diagonal GO methods. Numer. Math. 94(1), 93--106 (2003)

### 2019-08-27 2019/NUMTA_Barkalov/LNCS/numta19.tex
a point is corrected
Fix some minor comments from reviewers

@@
 %
-\author{Roman Strongin%\orcidID{0000-0003-0390-6695} 
-\and Konstantin Barkalov%\orcidID{0000-0001-5273-2471} 
-\and Semen Bevzuk%\orcidID{0000-0002-3845-5356}
+\author{Roman Strongin\orcidID{0000-0003-0390-6695} 
+\and Konstantin Barkalov\orcidID{0000-0001-5273-2471} 
+\and Semen Bevzuk\orcidID{0000-0002-3845-5356}
 }
@@
 \end{gather}
-where the objective function is a black-box function and assumed to satisfy the Lipschitz condition
+where the objective function is a black-box function and it is assumed to satisfy the Lipschitz condition
 \[
@@
 \]
-with the constant $L$ unknown a priori.
+with the constant $L, \; L<\infty,$ unknown a priori.
 
@@
 The assumption of the objective function to be Lipschitzian is typical for many approaches to the development of the global optimization algorithms \cite{Evtushenko2013,Zilinskas2010,Pinter1996,Strongin2000}.
-At that, the adaptive estimate of the unknown Lipschitz constant based on the obtained search information is the most important problem being solved in these algorithms anyway.
-The Lipschitz constant value affects the convergence rate of the global optimization algorithms essentially. Therefore, the issue of its correct estimate is so important. 
-The underestimate of the constant true value may result in losing the convergence of the algorithm to the global solution.
-At the same time, too large value of the estimate of the constant for the objective function not matching its real behavior results in slow convergence of the algorithm to the global minimizer.
+Moreover, the adaptive estimate of the unknown Lipschitz constant, based on the obtained search information, is one of the most important problems being solved in these algorithms.
+The value of the Lipschitz constant affects the convergence rate of the global optimization algorithms essentially. Therefore, the issue of its correct estimate is so important. 
+The underestimation of the real value of this constant may result in losing the convergence of the algorithm to the global solution.
+At the same time, a too large value of the estimate of the constant could bring the objective function not matching its real behavior.
+This produces slow convergence of the algorithm to the global minimizer.
 
@@
 For example, the use of the global estimate over the whole search domain can slow down the convergence of the algorithm to the global minimizer.
-The use of the local estimates accelerating the convergence requires an adequate adjustment of the algorithm parameters in order to preserve the global convergence.
+The use of the local estimates to accelerate the convergence of the method requires an adequate adjustment of the parameters of the algorithm in order to preserve the global convergence.
 
-In the present work, an algorithm, in which it is proposed to use two global estimates of the Lipschitz constant, one being much greater than another, is considered. 
+In the present work, it is considered a new algorithm in which there is the use of two global estimates of the Lipschitz constant. One of the two estimates is much greater than the other one. 
 The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
-The choice of the one of two estimates to be used in the algorithm rules is performed adaptively subject to the function behavior and to the search phase.
+The choice of one of the two estimates in the algorithm is performed adaptively during the search phase.
 
 The rigorous substantiation of the proposed approach goes beyond the present initial publication and will be done in the forthcoming works.
-Here we present the results of numerical experiments clearly demonstrating the efficiency of the novel algorithm.
-When conducting the numerical experiments, several hundred multiextremal test problems of various dimensionalities have been solved.
+Here we present the results of numerical experiments that clearly demonstrate the efficiency of the new algorithm.
+Several hundred multiextremal test problems of various dimensionalities have been solved conducting numerical experiments.
 
@@
 
-The adaptation of the efficient algorithms of solving the one-dimensional problems to solving the multidimensional problems is typical method of constructing the global optimization algorithms, see, for example, the diagonal partitions method in \cite{Sergeyev2006} or the simplicial partitions method in \cite{Zilinskas2008}.
+The adaptation of the efficient algorithms that solve one-dimensional problems to solve multidimensional problems is a typical method to construct global optimization algorithms, see, for example, the diagonal partitions method in \cite{Sergeyev2006} or the simplicial partitions method in \cite{Zilinskas2008}.
 
-In the present paper, the approach based on the ideas of dimensionality reduction with the use of the Peano-Hilbert curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously maps the unit interval $[0,1]$ onto the $n$-dimensional cube $D$ from (\ref{D}) has been used.
+In the present paper, the approach based on the ideas of reduction of dimension with the use of the Peano-Hilbert curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously maps the unit interval $[0,1]$ onto the $N$-dimensional cube $D$ from (\ref{D}) has been used.
 By using this kind of mapping, it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem
@@
 with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation
-$ H=2 L \sqrt{N+3}$.
+$ H=2 L \sqrt{N+3}$ and $y(x)$ is a Peano-Hilbert curve from $[0,1]$ onto $D$.
 
-Let us call the process of computing a function value (including the construction of the image $y=y(x)$) a \textit{trial}, and the pair $\{x, z = \varphi(y(x))\}$ the outcome of the trial.
+Let us call the process of computing a function value (including the construction of the image $y=y(x)$) as a \textit{trial}, and the pair $\{x, z = \varphi(y(x))\}$ as the outcome of the trial.
 
-The global search algorithm (according to \cite{Strongin2000}) can be formulated as follows.
+The Divide-The-Best global search algorithm used in this paper (according to \cite{Strongin2000}) can be formulated as follows.
 The first two trials are executed at 
@@
 Compute the maximum absolute value of the first divided differences
+\[
+M = \max_{1 \leq i \leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
+\]
+where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$ and let
 \begin{equation}\label{mu}
-\mu = \max_{1 \leq i \leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
+\mu = \left\{
+   \begin{array}{lr}
+     1, &\textrm{if } M = 0,\\
+     M, &\textrm{if } M \neq 0.
+   \end{array}
+\right.
 \end{equation}
-where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$. If $\mu = 0$, set $\mu = 1$.
+
+ %If $\mu = 0$, set $\mu = 1$.
 \item
@@
 The theory of convergence of this algorithm is provided in \cite{Strongin2000}. 
-The algorithm is allows for an efficient
-parallelization for shared and distributed memory \cite{Gergel2003}  and for accelerators \cite{Gergel2016}.
+The algorithm can be efficiently parallelized for shared and distributed memory \cite{Gergel2003}  and for accelerators \cite{Gergel2016}.
 
@@
 \noindent is satisfied.
-Thus,  an appropriate choice of the parameter $r$ from (\ref{R}) allows using the value $r\mu / 2^{3-1/N}\sqrt{N+3}$ as an estimate of the Lipschitz constant for the objective function $\varphi(y)$.
+Thus,  an appropriate choice of the parameter $r$ from (\ref{R}) allows using the value $(r\mu) / (2^{3-1/N}\sqrt{N+3})$ as an estimate of the Lipschitz constant for the objective function $\varphi(y)$.
 
@@
 However, in this case the method will perform a large number of trials until the stop condition is satisfied.
-The choice of small value of the parameter $r$ (that corresponds to the lower estimate of the Lipschitz constant) would considerably reduce the number of trials but may violate the convergence to the global extremum.
+The choice of a small value of the parameter $r$ (that corresponds to the lower estimate of the Lipschitz constant) would considerably reduce the number of trials but may violate the convergence to the global extremum.
 
 An approach, in which two estimates of the Lipschitz constant %, one of which is essentially greater than another, 
-are used in the rules of algorithm, is a promising one. 
-This approach implies the use of two parameters $r_{glob}$ and $r_{loc}$ where $r_{glob} > r_{loc}$ in the algorithm.
+are used in the rules of the algorithm, is a promising one. 
+This approach implies the use of two parameters $r_{glob}$ and $r_{loc}$ where $r_{glob} > r_{loc}>1$ in the algorithm.
 
-The rules of the algorithm with two estimates of the Lipschitz constant reproduce the ones of global search algorithm completely except Rule 3 (computing the characteristic).
+The rules of the algorithm with two estimates of the Lipschitz constant reproduce the ones of global search algorithm completely except Rule 3 (the computation of the characteristic).
 
@@
 This value of $r$ is used in Rule 5 of the algorithm in the computing of the next trial point.   
-\end{itemize}.
+\end{itemize}
 
 This method of computing the characteristic is substantiated as follows. 
-Let current minimum value of $z^*$ from (\ref{z}) is achieved at the left point of the $i^{\rm th}$ interval, i.e.
+Let current minimum value of $z^*$ from (\ref{z}) be achieved at the left point of the $i^{\rm th}$ interval, i.e.
 $z^* = z_{i-1}$.
@@
 optimization algorithms ~\cite{Barkalov2018,Paulavicius2014,Sergeyev2015}.
-Six GKLS classes of differentiable test functions of the dimensions $N = 3,4,5$
+Six GKLS classes of 100 differentiable test functions of the dimensions $N = 3,4,5$
 have been used. For each dimension, both \textit{Hard} and \textit{Simple}
@@
 algorithm generated a trial point $y^k$ in the vicinity of the global minimum,
-i.e. $\left|y^k-y^\ast\right| <\delta\left\|b-a\right\|$, 
+i.e. $\left\|y^k-y^\ast\right\| <\delta\left\|b-a\right\|$, 
 $a$ and $b$ are the boundaries of the search domain $D$.

### 2019-08-28 2019/NUMTA_Barkalov/LNCS/numta19.tex
Some major correction in Russian

@@
 
-%\usepackage[utf8x]{inputenc}
-%\usepackage[english,russian]{babel}
+\usepackage[utf8x]{inputenc}
+\usepackage[english,russian]{babel}
 
@@
  \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problem}\\
- D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\} \label{D},
+ D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \; a_i,b_i\in R, \;  1\leq i \leq N\right\} \label{D},
 \end{gather}
@@
 
-%Russian
 The assumption of the objective function to be Lipschitzian is typical for many approaches to the development of the global optimization algorithms \cite{Evtushenko2013,Zilinskas2010,Pinter1996,Strongin2000}.
@@
 \section{Global search algorithm and dimensionality reduction}
-%Russian
 
@@
 $ H=2 L \sqrt{N+3}$ and $y(x)$ is a Peano-Hilbert curve from $[0,1]$ onto $D$.
+%Russian
+Отметим, что теоретически Peano-Hilbert curve $y(x)$ определяется as a limit object.
+Поэтому при практической реализации может быть построено лишь некоторое приближение к истинной space-filling curve.
+Методы построения подобных approximations (называемых evolvents) рассмотрены в \cite{Sergeyev2013,Strongin2000}.
+При этом точность приближения развертки к истинной кривой $y(x)$ зависит от плотности развертки $m$ (являющейся параметром ее построения) и составляет величину порядка $2^{-m}$ по каждой координате.
 
@@
 
- %If $\mu = 0$, set $\mu = 1$.
 \item
@@
 
-%Russian
 The global search algorithm presented in previous section is intended for solving the multiextremal problems, in which the objective function satisfies the Lipschitz condition.
@@
 
-The rules of the algorithm with two estimates of the Lipschitz constant reproduce the ones of global search algorithm completely except Rule 3 (the computation of the characteristic).
+%Russian
+The rules of the algorithm with two estimates of the Lipschitz constant reproduce the ones of global search algorithm completely except Rule 3 (the computation of the characteristic) and Rule 4 (поиск интервала с максимальной характеристикой).
 
@@
 \end{equation}   
-and fix the value $r = r_{loc}$ if $\rho R_{loc}(i) > R_{glob}(i)$, otherwise fix $r=r_{glob}$.
+\end{itemize}
+
+%Russian
+Новые правила поиска интервала с максимальной характеристикой будут состоять в следующем:
+\begin{itemize}
+\item 
+Select the interval $(x_{t-1},x_t)$ corresponding to the maximum characteristic
+\[
+R(t)= \max_{1 \leq i \leq k}R(i).
+\]
+\item
+Fix the value $r = r_{loc}$ if $\rho R_{loc}(t) > R_{glob}(t)$, otherwise fix $r=r_{glob}$.
 \item
@@
 
-The numerical comparison of the algorithms has been carried out using the GKLS 
-test problem generator \cite{Gaviano2003}. 
-This generator of multiextremal functions is often used for the investigations of the global 
-optimization algorithms ~\cite{Barkalov2018,Paulavicius2014,Sergeyev2015}.
-Six GKLS classes of 100 differentiable test functions of the dimensions $N = 3,4,5$
-have been used. For each dimension, both \textit{Hard} and \textit{Simple}
-classes have been considered. The difficulty of a class was increased either by
-decreasing the radius of the attraction region of the global minimizer, or by
-decreasing the distance from the global minimizer $y^\ast$ to the domain
-boundaries. The global minimizer $y^\ast$ was considered to be found, if the
-algorithm generated a trial point $y^k$ in the vicinity of the global minimum,
-i.e. $\left\|y^k-y^\ast\right\| <\delta\left\|b-a\right\|$, 
-$a$ and $b$ are the boundaries of the search domain $D$.
+%Russian
+Численное сравление алгоритмов было выполнено посредством решения нескольких серий задач, порожденных генератором GKLS \cite{Gaviano2003}.
+Данный генератор многоэкстремальных функций широко используется для сравнения методов глобальной оптимизации (see, for example, \cite{Barkalov2018,Paulavicius2014,Sergeyev2015}).
+В данном исследовании были решены шесть серий по 100 задач размерностей $N = 3,4,5$.
+Для каждой размерности генерировались \textit{Simple} and \textit{Hard} problems, отличающиеся размерами attraction region локальных и глобального minimizer.
+Задача считалась решенной, если метод проводил очередное испытание в точке $y^k$, которая находилась в $\delta$-октрестности global minimizer $y^\ast$, 
+т.е. $\left\|y^k-y^\ast\right\| <\delta\left\|b-a\right\|$, где  $a$ and $b$ --- границы search domain $D$.
 
-%Russian
+The numerical comparison of the algorithms has been carried out using the GKLS test problem generator \cite{Gaviano2003}.
+This generator of multiextremal functions is often used for the investigations of the global optimization algorithms ~\cite{Barkalov2018,Paulavicius2014,Sergeyev2015}.
+Six GKLS classes of 100 differentiable test functions of the dimensions $N = 3,4,5$ have been used.
+For each dimension, both \textit{Hard} and \textit{Simple} classes have been considered.
+The difficulty of a class was increased either by decreasing the radius of the attraction region of the global minimizer, or by decreasing the distance from the global minimizer $y^\ast$ to the domain boundaries.
+The global minimizer $y^\ast$ was considered to be found, if the algorithm generated a trial point $y^k$ in the vicinity of the global minimum, i.e.
+$\left\|y^k-y^\ast\right\| <\delta\left\|b-a\right\|$, $a$ and $b$ are the boundaries of the search domain $D$.
+
 Each series of problems has been solved by the original global search algorithm (GSA) and by the method with two estimates of the Lipschitz constant (GSA-DL).
 The evolvent constructed using the parameter $m = 10$ was used for the dimensionality reduction in both algorithms.
-The relative accuracy of the solution search was $\delta = 0.01$.% for $N=3,4$ and $\delta = 0.02$ for $N=5$.
+The relative accuracy of the solution search was $\delta = 0.01$.
 The maximum allowable number of iterations per problem was $K_{max} = 10^6$.
@@
 
-The advantage of the GSA-DL algorithm over its prototype is confirmed by the operational characteristics of the algorithms as well.
-
-Assume a series of test problems to be solved.
-The results of solving the series can be presented by a function $p(k)$ featuring the fraction of the total number of problems solved in $k$ iterations.
-Such a function will be called the \textit{operational characteristic} of the algorithm.
-
-The operational characteristics for the GSA and GSA-DL methods obtained when solving the \textit{Simple} and \textit{Hard} problem series with the dimensionalities $N=4$ and $N=5$ are presented in Fig.
-\ref{oper4} and \ref{oper5}, respectively.
-The values of the parameters $r$ used for estimating the Lipschitz constant are given in the figures.
-
 \begin{figure}
@@
 
+The advantage of the GSA-DL algorithm over its prototype is confirmed by the operational characteristics of the algorithms as well.
+
+Assume a series of test problems to be solved.
+The results of solving the series can be presented by a function $p(k)$ featuring the fraction of the total number of problems solved in $k$ iterations.
+Such a function will be called the \textit{operational characteristic} of the algorithm.
+
+The operational characteristics for the GSA and GSA-DL methods obtained when solving the \textit{Simple} and \textit{Hard} problem series with the dimensionalities $N=4$ and $N=5$ are presented in Fig.
+\ref{oper4} and \ref{oper5}, respectively.
+The values of the parameters $r$ used for estimating the Lipschitz constant are given in the figures.
+
 The lower curves in Fig.
@@
 Such positions of the curves show the algorithm with two estimates of the Lipschitz constant provides much faster solving of the problem series in average that the algorithm using single estimate of the constant.
+%Russian
+Отметим, что для решения задач всех классов (кроме класса  \textit{Hard} при $N=5$) метод GSA-DL требует примерно в два раза меньше испытаний, чем GSA.
+Ухудшение результатов на классе \textit{Hard} при $N=5$ объясняется сложностью функций данного класса, характеризующихся большими attraction region локальных минимумов и узкой attraction region глобального минимума.
+Для корректного решения таких задач метод  GSA-DL чаще использует бОльшую оценку константы Липшица, что сближает скорость работы алгоритмов GSA и GSA-DL.
 

### 2019-08-29 2019/NUMTA_Barkalov/LNCS/numta19.tex
Correct all reviewers comments in Russian

@@
 This method of computing the characteristic is substantiated as follows. 
+%Russian
+На каждой итерации будет возникать интервал, на одной из границ которого достигается текущее минимальное значение $z^*$ from (\ref{z}).
+Причем на заключительной фазе поиска данный интервал будет соответствовать интервалу, содержащему глобальный минимум, т.е.
+будет являться лучшим с точки зрения проведения дальнейших испытаний в нем.
+
 Let current minimum value of $z^*$ from (\ref{z}) be achieved at the left point of the $i^{\rm th}$ interval, i.e.
@@
 \]
-Consequently, two lower estimates of the two characteristics of current best interval $R_{loc}(i)$ and $R_{glob}(i)$ calculated with different parameters $r_{loc}$ and $r_{glob}$ will take different values.
+Следовательно, для нижних оценок характеристик $R_{loc}(i)$ and $R_{glob}(i)$ calculated with different parameters $r_{loc}$ and $r_{glob}$, будет выполняться соотношение
+\[
+	\Delta_i \left( 1 - \frac{1}{r_{glob}} \right)^2 > \Delta_i \left( 1 - \frac{1}{r_{loc}} \right)^2.
+\]
+
+Таким образом, если выбирать наибольую из характеристик в соответствии с
+\[
+R(i) = \max\{R_{loc}(i),R_{glob}(i)\}
+\]
+то, как правило, будет выбираться характеристика $R_{glob}(i)$, соответствующая большей оценке Lipschitz constant; меньшая оценка (ускоряющая процесс уточнения текущего решения) использоваться не будет.
+Однако если домножить нижнюю оценку для характеристики $R_{loc}(i)$ на коэффициент $\rho$ в соответствии с (\ref{pho}), то указанные нижние оценки будут равны, что сделает более вероятным выбор характеристики $R_{loc}(i)$, соответствующей меньшей оценке Lipschitz constant.
+
+%Consequently, two lower estimates of the two characteristics of current best interval $R_{loc}(i)$ and $R_{glob}(i)$ calculated with different parameters $r_{loc}$ and $r_{glob}$ will take different values.
 However, if the former is multiplied by the coefficient $\rho$ from (\ref{pho}), the lower estimates of the characteristics of the best interval would match each other that would make the characteristics calculated for other intervals comparable as well.
@@
 Численное сравление алгоритмов было выполнено посредством решения нескольких серий задач, порожденных генератором GKLS \cite{Gaviano2003}.
-Данный генератор многоэкстремальных функций широко используется для сравнения методов глобальной оптимизации (see, for example, \cite{Barkalov2018,Paulavicius2014,Sergeyev2015}).
+Данный генератор of multiextremal functions широко используется для сравнения методов глобальной оптимизации (see, for example, \cite{Barkalov2018,Paulavicius2014,Sergeyev2015}).
 В данном исследовании были решены шесть серий по 100 задач размерностей $N = 3,4,5$.
-Для каждой размерности генерировались \textit{Simple} and \textit{Hard} problems, отличающиеся размерами attraction region локальных и глобального minimizer.
+Для каждой размерности генерировались \textit{Simple} and \textit{Hard} problems, отличающиеся размерами attraction region локальных и глобального экстремума.
 Задача считалась решенной, если метод проводил очередное испытание в точке $y^k$, которая находилась в $\delta$-октрестности global minimizer $y^\ast$, 
-т.е. $\left\|y^k-y^\ast\right\| <\delta\left\|b-a\right\|$, где  $a$ and $b$ --- границы search domain $D$.
+т.е. $\left\|y^k-y^\ast\right\| <\delta\left\|b-a\right\|$, where  $a$ and $b$ --- границы search domain $D$.
 
-The numerical comparison of the algorithms has been carried out using the GKLS test problem generator \cite{Gaviano2003}.
+%The numerical comparison of the algorithms has been carried out using the GKLS test problem generator \cite{Gaviano2003}.
 This generator of multiextremal functions is often used for the investigations of the global optimization algorithms ~\cite{Barkalov2018,Paulavicius2014,Sergeyev2015}.

### 2019-08-29 2019/Automation/Barkalov.tex
First full version

@@
 \documentclass[12pt]{a&t}
+\usepackage{hyperref}
 \usepackage{graphicx}
@@
 \year{2019}
-\title{НАЗВАНИЕ СТАТЬИ ИЛИ ЗАМЕТКИ БЫВАЕТ С ФОРМУЛАМИ $a+b=c$}%
+\title{АДАПТИВНАЯ ГЛОБАЛЬНАЯ ОПТИМИЗАЦИИ НА ОСНОВЕ БЛОЧНО-РЕКУРСИВНОЙ СХЕМЫ РЕДУКЦИИ РАЗМЕРНОСТИ}%
 \thanks{Исследование выполнено за счет гранта Российского научного фонда (проект \mbox{№\,16-11-10150}).}
@@
 \begin{abstract}
-Краткая аннотация статьи или заметки. Иногда не бывает. Хх
-хххххххххххх хххххх хххххххх х хххххххххххххх ххх ххххххх хххх.
-Хххххххххххххх ххх ххххх хххххх х хххххххххххххх ххх ххххххх
-хххх.
+В статье рассматриваются задачи многомерной многоэкстремальной оптимизации и численные методы их решения.
+Об оптимизируемой функции делается лишь общее предположение, что она удовлетворяет условию Липшица с априори неизвестной константой.
+Задачи такого типа часто встречаются в приложениях.
+Рассмотрено два способа редукции размерности в задачах многомерной оптимизации: ипользование кривых Пеано (разверток) и рекурсиваная многошаговая схема.
+Предложена обобщенная схема, комбинирующая эти два подхода.
+В новой схеме решение многомерной задачи сводится к решению семейства задач меньшей размерности, в которых, в свою очередь, используются развертки.
+Реализован адаптивный алгоритм, в котором все возникающие подзадачи решаются одновременно.
+Проведены численные эксперименты на нескольких сотнях тестовых задач, подтверждающие эффективность предложенной схемы редукции размерности.
 \end{abstract}
@@
 Неймарком \cite{Neymark66a,Neymark66b} и развиты Р.Г.
-Стронгиным \cite{Strongin78}.
+Стронгиным \cite{Strongin70,Strongin78}.
 Впоследствии метод, изначально предложенный для решения безусловных задач, был успешно обобщен для решения задач с невыпуклыми ограничениями \cite{Strongin86,Strongin87} и многокритериальных задач \cite{Strongin93}.
+Для различных вариантов алгоритма были предложены способы распараллеливания, учитывающие особенности архитектуры современных вычислительных систем \cite{Strongin13}.
 
@@
 \section{Базовый алгоритм глобального поиска}
+\label{SectionCore}
 
-\section{Многошаговая схема редукции размерности}
+В качестве базовой задачи рассмотрим одномерную задачу многоэкстремальной оптимизации 
+\begin{equation}\label{uni_problem}
+\varphi^\ast = \varphi(x^\ast)=\min{\left\{\varphi(x):x\in \left[0,1\right] 
+\right\}}
+\end{equation}
+с целевой функцией, удовлетворяющей условию Липшица
 
-Многошаговая схема редукции размерности (схема вложненной оптимизации) основана на известном соотношении
+Приведем описание алгоритма глобального поиска (АГП) в соответствии с \cite{Strongin00}.
+В процессе своей работы АГП породждает последовательность точек $x^i$, в которых вычисляются значения целевой функции $z^i=\varphi(x^i)$. 
+Будем называть процесс вычисления значения целевой функции \textit{испытанием}.
+
+В соответствии с алгоритмом, первые два испытания проводятся в граничных точках интервала $[a,b]$, т.е., $x^0=a,\;x^1=b$. 
+В этих точка вычисляются значения целевой функции $z^0=\varphi(x^0),\;z^1=\varphi(x^1)$, и устанавливается значение счетчика $k=1$. 
+Точка очередного испытания $x^{k+1}, k\geq 1,$ выбирается в соответствии со следующими правилами.
+
+\textbf{Шаг 1.} Перенумеровать нижним индексом (начиная с 0) точки $x^i,\:0\leq i\leq k$, проведенных испытаний в порядке возрастания координаты, т.е.,
+\[
+0=x_0<x_1<\ldots <x_{k}=1.
+\] 
+Сопоставить точкам $x_i, 0\leq i\leq k$, вычисленные в них значения целевой функции $z_i=\varphi(x_i), 0\leq i\leq k$.
+
+\textbf{Шаг 2.} Вычислить максимальное абсолютное значение относительной первой разности
+\begin{equation}\label{mu}
+\mu=\max_{1\leq i\leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i}
+\end{equation}
+где $\Delta_i = x_i-x_{i-1}$. Если вычисленное в соответствии с данной формулой значение равно 0, то положить $\mu = 1$.
+
+\textbf{Шаг 3.} Для всех интервалов $(x_{i-1},x_i),1\leq i\leq k$,  вычислить значение
+\begin{equation}\label{R}
+R(i)=r\mu\Delta_i+\frac{(z_i-z_{i-1})^2}{r\mu\Delta_i}-2(z_i+z_{i-1}),
+\end{equation} 
+называемое \textit{характеристикой} интервала; величина $r>1$ является параметром алгоритма.
+
+\textbf{Шаг 4.} Найти интервал $(x_{t-1},x_t)$ с максимальной характеристикой
+\begin{equation}\label{MaxR}
+R(t)=\max_{1\leq i\leq {k}}R(i).
+\end{equation}.
+Если максимальная характеристика соответствует нескольким интервалам, то в качестве $t$ выбрать минимальное число, удовлетворяющее (\ref{MaxR}).
+
+\textbf{Шаг 5.} Провести новое испытание в точке
+\begin{equation}\label{xk1}
+x^{k+1}=\frac{1}{2}(x_{t-1}+x_t) - \frac{z_t-z_{t-1}}{2r\mu}.
+\end{equation}
+
+Алгоритм прекращает свою работу при выполнении условия $\Delta_t<\epsilon$; здесь $t$ из (\ref{MaxR}), а $\epsilon>0$ --- заданная точность. 
+В качестве оценки решения задачи выбираются значения 
+\[
+z_k^\ast=\min_{0\leq i \leq k}\varphi(x^i), \ x_k^\ast=\arg \min_{0\leq i \leq
+ k}\varphi(x^i).
+\] 
+Теоретические условия, определяющие сходимость алгоритма, представлены в \cite{Strongin00}.
+
+\section{Редукция размерности}
+\subsection{Редукция размерности с помощью кривых, заполняющих пространство}
+\label{SectionPeano}
+
+Рассмотрим схему редукции размерности, основанную на использовании кривых, заполняющих пространство (кривых Пеано).
+Известно, что подобного типа кривые позволяют однозначно отобразить одномерный отрезок $[0,1]$ на многомерную область, т.е.
+\begin{equation}
+\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq N\right\}=\left\{
+y(x):0\leq x \leq 1 \right\}
+\end{equation}
+
+Отметим, что теоретически кривая $y(x)$ определяется как предельнгый объект.
+Поэтому при практической реализации может быть построено лишь некоторое приближение к истинной кривой.
+Методы построения подобных приближений (называемых \textit{развертками}) рассмотрены в \cite{Strongin78,Strongin00}.
+При этом точность приближения развертки к истинной кривой $y(x)$ зависит от \textit{плотности развертки} $m$ (являющейся параметром ее построения) и составляет величину порядка $2^{-m}$ по каждой координате.
+
+Используя подобного род отображения можно свести решение многомерной задачи (\ref{main_problem}) к решению эквивалетной ей одномерной
+\begin{equation}
+\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]
+\right\}}.
+\end{equation}
+
+Важным свойством при этом является сохранение ограниченности относительных разностей функции 
+(см. \cite{Strongin00}). Если функция $\varphi(y)$ в области $D$ удовлетворяла условию Липшица, тогда редуцированная функция  
+$\varphi(y(x)), x\in [0,1]$ будет удовлетворять равномерному условию Гельдера 
+\begin{equation}\label{Holder}
+\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N},
+\end{equation}
+где константа Гельдера $H$ связана с константой Липшица  $L$ соотношением
+\begin{equation}\label{HL}
+H=2L\sqrt{N+3}.
+\end{equation}
+
+Условия (\ref{Holder}), (\ref{HL}) позволяет легко обобщить <<одномерный>> алгорим из раздела \ref{SectionCore} для решения многомерных задач. 
+Для этого достаточно заменить длины интервалов $\Delta_i$ в формулах (\ref{mu}),(\ref{R}) на длины, вычисляемые в соответствии с 
+\begin{equation}
+\Delta_i=\left(x_i-x_{i-1}\right)^{1/N},
+\end{equation}
+а также заменить формулу (\ref{xk1}) для вычисления точки нового испытания на формулу
+\begin{equation}
+x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}
+\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N.
+\end{equation}
+
+\subsection{Многошаговая схема редукции размерности}
+
+Многошаговая схема редукции размерности (схема вложненной оптимизации) основана на известном соотношении %ссылка!!!! Carr? 
 \begin{equation}\label{nested}
@@
 
-Дадим более подробное описание взаимосвязей между одномерными подзадачами в многошаговой схеме. Для удобства последующего изложения введем обозначение
-\[
-v_i = (y_1,...y_i), \; 1 \leq i\ \leq N.
-\]
-Тогда одномерные функции $\overline{\varphi}_i(y_i), \; 1 < i\ \leq N$, порожденные в рамках многошаговой схемы, могут быть представлены в виде
-\[
-\overline{\varphi_i}(y)_i = \varphi_i(v_{i-1},y_i), \; 1 < i\ \leq N,
-\]
-где вектор $v_{i-1}$ фиксирован.
+%Дадим более подробное описание взаимосвязей между одномерными подзадачами в многошаговой схеме. %Для удобства последующего изложения введем обозначение
+%\[ v_i = (y_1,...y_i), \; 1 \leq i\ \leq N. \]
+%Тогда одномерные функции $\overline{\varphi}_i(y_i), \; 1 < i\ \leq N$, порожденные в рамках многошаговой схемы, могут быть представлены в виде
+%\[ \overline{\varphi_i}(y)_i = \varphi_i(v_{i-1},y_i), \; 1 < i\ \leq N, \]
+%где вектор $v_{i-1}$ фиксирован.
 
-В соответствии с многошаговой схемой на некотором шаге поиска $i, 1 \leq i \leq k_1$  алгоритм начинает минимизировать одномерную функцию $\varphi_1(y_1)$.
+%В соответствии с многошаговой схемой на некотором шаге поиска $i, 1 \leq i \leq k_1$  алгоритм начинает минимизировать одномерную функцию $\varphi_1(y_1)$.
 Для этого, используя накопленную поисковую информацию, алгоритм выбирает точку очередной итерации $\widehat{y}_1$ и вычисляет значение функции в ней.
@@
 Данный процесс выполняется рекурсивно до последнего $N$-го уровня.
-Возникающие при этом подзадачи и взаимосвязи между ними отображены на рис.
-\ref{fig0}.
+
+%А нужен ли рисунок?
+Возникающие при этом подзадачи и взаимосвязи между ними отображены на рис. \ref{fig0}.
+Наглядно видно, что структура взаимосвязей имеет форму дерева.
+Указанная структура может меняться в процессе решения многомерной задачи (\ref{main_problem}): вычисление значения функции $\varphi_i(\widehat{v}_{i-1},y_i), \; 1\leq i < N,$  на $i$-м уровне рекурсии требует решения всех подзадач в одном из поддеревьев $(i+1)$-го уровня.
+При этом функции $\varphi_N(\widehat{v}_{N-1},y_N)$ на $N$-м уровне являются листьями дерева задач, их значения вычисляются непосредственно.
 
@@
 
-Наглядно видно, что структура взаимосвязей имеет форму дерева.
-Указанная структура может меняться в процессе решения многомерной задачи (\ref{main_problem}): вычисление значения функции $\varphi_i(\widehat{v}_{i-1},y_i), \; 1\leq i < N,$  на $i$-м уровне рекурсии требует решения всех подзадач в одном из поддеревьев $(i+1)$-го уровня.
-При этом функции $\varphi_N(\widehat{v}_{N-1},y_N)$ на $N$-м уровне являются листьями дерева задач, их значения вычисляются непосредственно.
+Для описанной выше многошаговой схемы было предложено обобщение --- \textit{блочная многошаговая схема}, которое комбинирует использование разверток и вложенной оптмизации \cite{Barkalov2014}.
 
-%недостатки многошаговой схемы
+Рассмотрим $y$ как вектор, состоящий из векторов (блочных переменных)
+\begin{equation}
+y=(y_1,y_2,...,y_N)=(u_1,u_2,...,u_M),
+\end{equation}
+где $i$-я блочная переменная, т.е. вектор $u_i$, состоит из взятых последовательно компонент вектора $y$, т.е. $u_1=(y_1,y_2,...,y_{N_1})$, $u_2=(y_{N_1+1},y_{N_1+2}
+,...,y_{N_1+N_2})$,..., $u_M=(y_{N-N_M+1},y_{N-N_M+2},...,y_{N})$, а $N_1+
+N_2+...+N_M=N$.
 
-\section{Адаптивная многошаговая схема}
+Используя введенные обозначения, основное соотношение многошаговой схемы (\ref{nested}) 
+может быть переписано в форме 
+\begin{equation}\label{block_nested}
+\min_{y \in D}\varphi(y) = \min_{u_1\in D_1}\min_{u_2\in D_2}...\min_{u_M \in D_M}\varphi(y),
+\end{equation}
+где подобласти $D_i, 1 \leq i \leq M$, являются проекциями исходной области поиска 
+$D$ на подпространства, соответствующие переменным $u_i, 1 \leq i \leq M$.
 
-Рассмотрим обобщение схемы вложенной оптимизации (\ref{nested_N})--(\ref{nested_1}), позволяющее преодолеть недостатки указанные выше.
-Основным отличием является отказ от принципа подчитенности одномерных подзадач на основе функций $\overline{\varphi_i}(y_i), \; 1\leq i \leq N,$ порожденных в рамках многошаговой схемы.
-В новом подходе все возникающие подзадачи предлагается решать одновременно.
-Предварительное описание \textit{адаптивной многошаговой схемы} редукции размерности заключается в следующем.
+Способ решения задачи (\ref{main_problem}) на основе соотношения (\ref{block_nested}) будет в целом повторять  (\ref{nested_N})--(\ref{nested_1}).
+Необходимо лишь заменить исходные переменны $y_i, 1\leq i \leq N$,  на блочные переменные variables $u_i, 1\leq i \leq M$.
 
-%взять первый вариант из статьи в JoGO
+При этом вложенные подзадачи
+\begin{equation}\label{block_nested_i}
+\varphi^i(u_1,...,u_i) = \min_{u^{i+1}\in D_{i+1}} \varphi_{i+1}(u_1,...,u_i,u
+_{i+1}), 1\leq i\leq M-1.
+\end{equation}
+будут многомерными, что является основным отличием от исходной многошаговой схемы. Для решения многомерных подзадач может быть использована схема редукции размерности на основе кривых Пеано.
+
+\subsection{Блочная адаптивная схема редукции размерности}
+\Russian
+Решение возникающего множества подзадач вида (\ref{nested_i}) (для многошаговой схемы) или вида (\ref{block_nested_i}) (для блочной многошаговой схемы) может быть организовано различными способами. 
+Очевидный способ (детально проработанный в \cite{Grishagin97,Grishagin2001,Grishagin2015} для многошаговой схемы и в \cite{Barkalov2014,Barkalov2016} для блочной многошаговой схемы) основан на решении поздадач в соответствии с рекурсивным порядком их порождения.
+Однако здесь возникает потеря значительной части информации о целевой функции.
+
+Иным  подходом является адаптивная схема, в которой все подзадачи решаются одновременно, что позволяет более полно учитывать информацию о многомерной задаче и за счет этого ускорять процесс ее решения.
+Для случая одномерных поздадач данный подход был теоретически обоснован и апробирован в \cite{Grishagin16,Grishagin2016_1,Grishagin18}.
+Настоящая работа предлагает обобщение адаптивной схемы для случая многомерных подзадач.
+Дадим краткое описание ее основных элементов.
+
+Пусть вложенные подзадачи вида (\ref{block_nested_i}) решаются с помощью алгоритма глобального поиска, описанного в разделе \ref{SectionPeano}.
+Тогда каждой подзадаче (\ref{block_nested_i}) можно присвоить некоторое числовое значение, называемое характеристикой этой задачи.
+К качестве такой характеристики можно взять значение $R(t)$ из (\ref{MaxR}), т.е.
+максимальную характеристику интервалов, сформированных в данной задаче.
+В соответствии с правилом вычисления характеристик (\ref{R}), чем выше значение характеристики, тем более перспективной является подзадача для продолжения поиска в ней глобального минимума исходной задачи (\ref{main_problem}).
+Поэтому на каждой итерации выбирается подзадача с максимальной характеристикой для проведения в ней очередного испытания.
+Это испытание либо вычисляет значение целевой функции $\varphi(y)$ (если выбранная подзадача принадлежала уровню $j=M$), либо порождает новые подзадачи согласно (\ref{block_nested_i}) при $j\leq M-1$.
+В последнем случае новые порожденные задачи добавляются к текущему множеству задач, вычисляются их характеристики и процесс повторяется.
+Завершение процесса оптимизации происходит, когда в корневой задаче выполняется условие остановки алгоритма, решающего эту задачу.
+
 \section{Результаты экспериментов}
@@
 
-В работах \cite{Grishagin18,Sovrasov19} было показано, что алгоритм глобального поиска (АГП) как с использованием разверток, так и в сочетании с адаптивной схемой редукции размерности превосходит многие известные алгоритмы глобальной оптимизации, включая методы DIRECT и DIRECT\textit{l}.
+В работах \cite{Grishagin18,Sovrasov19} было экспериментально установлено, что алгоритм глобального поиска (АГП) как с использованием разверток, так и в сочетании с адаптивной схемой редукции размерности превосходит многие известные алгоритмы глобальной оптимизации, включая методы DIRECT и DIRECT\textit{l}.
 Поэтому в данном исследовании мы ограничимся сравнением вариантов АГП в сочетании с различными схемами редукции размерности.
@@
 
+\bibitem{Strongin70}
+{\it Стронгин Р.Г.} Многоэкстремальная минимизация // Автомат. и телемех. 1970. № 7. С. 63--67.
+
 \bibitem{Strongin78}
@@
 
+\bibitem{Strongin13}
+{\it Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013.
+
 \bibitem{Strongin91}
@@
 
-\bibitem{Strongin13}
-{\it Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013.
-
 \bibitem{Grishagin97}
-
 {\it Grishagin V.A., Sergeyev Y.D., Strongin R.G.} Parallel characteristical algorithms for solving problems of global optimization // J. Glob. Optim. 1997. V. 10(2). P. 185--206.
@@
 {\it Gergel V., Grishagin V., Gergel A.} Adaptive nested optimization scheme for multidimensional global search // J. Glob. Optim. 2016. V. 66(1). P. 35--51.
+
+\bibitem{Barkalov2014}
+{\it Barkalov K., Gergel V.} Multilevel scheme of dimensionality reduction for parallel
+global search algorithms // In: OPT-i 2014 -- 1st International Conference on Engineering and Applied Sciences Optimization, Proceedings. 2014. P. 2111--2124.
+
+\bibitem{Grishagin2001}
+{\it Sergeyev Y., Grishagin V.} Parallel asynchronous global search and the nested
+optimization scheme // J. Comput. Anal. Appl. 2001. V. 3(2). P. 123--145.
+
+\bibitem{Grishagin2015}
+{\it Gergel V., Grishagin V., Israfilov R.} Local tuning in nested scheme of global
+optimization // Procedia Computer Science. 2015. V. 51(1). P. 865--874
+
+\bibitem{Barkalov2016}
+{\it Barkalov K., Lebedev I.} Solving multidimensional global optimization problems
+using graphics accelerators // CCIS. 2016. V. 687. P. 224--235.
+
+\bibitem{Grishagin2016_1}
+{\it Grishagin V., Israfilov R., Sergeyev Y.}: Comparative efficiency of dimensionality reduction schemes in global optimization // AIP Conference Proceedings. 2016. V. 1776, art. no. 060011.
+
+\bibitem{Grishagin18}
+{\it Grishagin V., Israfiov R., Sergeyev Y.} Convergence conditions and numerical comparison of global optimization methods based on dimensionality reduction schemes // Appl.
+Math.
+Comput.
+2018.
+V.
+318.
+P.
+270--280.
 
@@
 
-\bibitem{Grishagin18}
-{\it Grishagin V., Israfiov R., Sergeyev Y.} Convergence conditions and numerical comparison of global optimization methods based on dimensionality reduction schemes // Appl.
-Math.
-Comput.
-2018.
-V.
-318.
-P.
-270--280.
-
 \bibitem{Sovrasov19}
-{\it Sovrasov V.} Comparison of several stochastic and deterministic derivative-free global optimization algorithms // Lecture Notes in Computer Science. 2019. V. 11548. P. 70--81.
+{\it Sovrasov V.} Comparison of several stochastic and deterministic derivative-free global optimization algorithms // LNCS. 2019. V. 11548. P. 70--81.
 

### 2019-09-04 2019/Automation/Barkalov.tex
New figures

@@
 
-Схема вложенной оптмизации (\ref{nested_N})--(\ref{nested_1}) в предположении липшицевости одномерных функций $\overline{\varphi}_i(y_i)=\varphi_i(y_1,...,y_i)$ послужила основой для обобщения методов одномерной оптимизации [7,25,32,34,36 из JOGO] на случай многих переменных [3,12,25,26,36 из JOGO].
+%Схема вложенной оптмизации (\ref{nested_N})--(\ref{nested_1}) в предположении липшицевости одномерных функций $\overline{\varphi}_i(y_i)=\varphi_i(y_1,...,y_i)$ послужила основой для обобщения методов одномерной оптимизации [7,25,32,34,36 из JOGO] на случай многих переменных [3,12,25,26,36 из JOGO].
 
@@
 Возникающие при этом подзадачи и взаимосвязи между ними отображены на рис. \ref{fig0}.
-Наглядно видно, что структура взаимосвязей имеет форму дерева.
-Указанная структура может меняться в процессе решения многомерной задачи (\ref{main_problem}): вычисление значения функции $\varphi_i(\widehat{v}_{i-1},y_i), \; 1\leq i < N,$  на $i$-м уровне рекурсии требует решения всех подзадач в одном из поддеревьев $(i+1)$-го уровня.
-При этом функции $\varphi_N(\widehat{v}_{N-1},y_N)$ на $N$-м уровне являются листьями дерева задач, их значения вычисляются непосредственно.
+Наглядно видно, что структура взаимосвязей имеет форму дерева,
+%Указанная структура может меняться в процессе решения многомерной задачи (\ref{main_problem}): вычисление значения функции $\varphi_i(\widehat{v}_{i-1},y_i), \; 1\leq i < N,$  на $i$-м уровне рекурсии требует решения всех подзадач в одном из поддеревьев $(i+1)$-го уровня.
+причем функции $\varphi_N(y_1,...,y_N)$ на $N$-м уровне являются листьями дерева задач, их значения вычисляются непосредственно.
 
@@
 \begin{center}%[scale=0.3][width=0.75\textwidth]
-  \includegraphics[width=0.9\textwidth]{fig1.png} 
+  \includegraphics[width=0.7\textwidth]{fig1.pdf} 
   \caption{Взаимосвязи между подзадачами в многошаговой схеме}\label{fig0} 
@@
 Вторая серия экспериментов была проведена на четырехмерных задачах классов GKLS \textit{Simple} и GKLS \textit{Hard} (100 функций каждого класса).
-В табл.~\ref{tab2}  представлено среднее число испытаний, выполненных АГП с использованием адаптивной многошаговой схемы ($K_a$) и блочной адаптивной схемы ($K_ba$) с формированием двух уровней подзадач одинаковой размерности $N_1=N_2=2$.
+В табл.~\ref{tab2}  представлено среднее число испытаний, выполненных АГП с использованием адаптивной многошаговой схемы ($K_a$) и блочной адаптивной схемы ($K_{ba}$) с формированием двух уровней подзадач одинаковой размерности $N_1=N_2=2$.
 Отметим, что при использовании исходного варианта адаптивной многошаговой схемы при решении задачи размерности $N=4$ формируется четыре уровня одномерных подзадач, что усложняет их обработку.

### 2019-09-05 2019/Automation/Barkalov.tex
6d example

@@
 
-\authors{Р.Г.~СТРОНГИН, д-р~физ.-мат.~наук\\
+\authors{
+Р.Г.~СТРОНГИН, д-р~физ.-мат.~наук\\
 В.П.~ГГЕРГЕЛЬ, д-р~техн.~наук\\
@@
 
-Первая серия экспериментов была проведена на двумерных задачах классов GKLS \textit{Simple} и GKLS \textit{Hard} (100 функций каждого класса).
-В табл.~\ref{tab1}  представлено среднее число испытаний, выполненных АГП с использованием разверток ($K_e$), многошаговой схемы ($K_n$) и адаптивной многошаговой схемы ($K_a$).
-На рис.~\ref{fig2}(a,b) приведены операционные характеристики алгоритмов, полученные на указанных классах задач.
-Непрерывная линия соответствует алгоритму с использованием разверток, короткий пунктир --- адаптивной многошаговой схемы, длинный пунктир -- многошаговой схемы.
-Результаты экспериментов показывают, что АГП с использованием адаптивной многошаговой схемы показывает примерно одинаковое быстродействие  по сравнению с развертками, и оба они значительно превосходят алгоритм, использующий многошаговую схему.
-Поэтому в дальнейших экспериментах мы ограничимся сравнением различных вариантов адаптивной схемы редукции размерности.
-
 \begin{table}
 \centering
-\caption{Среднее число испытаний для двумерных задач}\label{tab1}
+\caption{Среднее число испытаний на классах GKLS \textit{Simple} и \textit{Hard}, $N=2$}\label{tab1}
 \begin{tabular}{lccc}
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \begin{minipage}{0.5\linewidth}
@@
 \end{minipage}
-\caption{Operating characteristics using 2d GKLS \textit{Simple} (a) and \textit{Hard} (b) classes}
+\caption{Операционные характеристики на классах GKLS \textit{Simple} (a) и \textit{Hard} (b), $N=2$}
 \label{fig2}
 \end{figure}
+
+Первая серия экспериментов была проведена на двумерных задачах классов GKLS \textit{Simple} и GKLS \textit{Hard} (100 функций каждого класса).
+В табл.~\ref{tab1}  представлено среднее число испытаний, выполненных АГП с использованием разверток ($K_e$), многошаговой схемы ($K_n$) и адаптивной многошаговой схемы ($K_a$).
+На рис.~\ref{fig2}(a,b) приведены операционные характеристики алгоритмов, полученные на указанных классах задач.
+Непрерывная линия соответствует алгоритму с использованием разверток, короткий пунктир --- адаптивной многошаговой схемы, длинный пунктир --- многошаговой схемы.
+Результаты экспериментов показывают, что АГП с использованием адаптивной многошаговой схемы показывает примерно одинаковое быстродействие  по сравнению с развертками, и оба они значительно превосходят алгоритм, использующий многошаговую схему.
+Поэтому в дальнейших экспериментах мы ограничимся сравнением различных вариантов адаптивной схемы редукции размерности.
 
@@
 
-На рас.~\ref{fig3}(a,b) приведены операционные характеристики алгоритмов, полученные на классах GKLS \textit{Simple} и GKLS \textit{Hard}.
+На рис.~\ref{fig3}(a,b) приведены операционные характеристики алгоритмов, полученные на классах GKLS \textit{Simple} и GKLS \textit{Hard}.
 Пунктирная линия соответствует алгоритму с использованием адаптивной, а непрерывная ---  блочной адаптивной схемы.
@@
 \centering
-\caption{Среднее число испытаний для четырехмерных задач}\label{tab2}
+\caption{Среднее число испытаний на классах GKLS \textit{Simple} и \textit{Hard}, $N=4$}\label{tab2}
 \begin{tabular}{lcc}
@@
 \end{minipage}
-\caption{Операционные характеристики на классах задач GKLS \textit{Simple} (a) и \textit{Hard} (b)}
+\caption{Операционные характеристики на классах GKLS \textit{Simple} (a) и \textit{Hard} (b), $N=4$}
 \label{fig3}
+\end{figure}
+
+Последняя серия экспериментов была проведена на шестимерных задачах класса GKLS \textit{Simple} (100 функций).
+Сравнивалась работа АГП с использованием разверток и блочной адаптивной схемы с формированием двух уровней подзадач одинаковой размерности $N_1=N_2=3$.
+Среднее число испытаний, выполненных АГП с использованием разверток, составило 102987, тогда как с использованием блочной адаптивной схемы  --- 75 390. 
+На рис.~\ref{fig4} приведены операционные характеристики алгоритмов. Пунктирная линия соответствует алгоритму с использованием разверток, а непрерывная --- блочной адаптивной схемы.
+
+\begin{figure}
+\center
+\includegraphics[width=0.5\linewidth]{6DSimple.pdf}
+\caption{Операционные характеристики на классе GKLS \textit{Simple}, $N=6$}
+\label{fig4}
 \end{figure}

### 2019-09-09 2019/NUMTA_Barkalov/LNCS/numta19.tex
Fix some mistakes

@@
 %
-\title{Acceleration of Global Search by Implementing Dual Estimates for Lipschitz Constant\thanks{This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.}}
+\title{Acceleration of Global Search by Implementing Dual Estimates for Lipschitz Constant\thanks{This research was supported by the Russian Science Foundation, project No.\,16-11-10150.}}
 %
@@
 class of problems require reliable estimates of the Lipschitz constant to be 
-introduced. The various approaches have been proposed to take into account both
+introduced. Various approaches have been proposed to take into account both
 global and local properties of the objective function. In particular, algorithms
@@
 Results of numerical experiments on the random sample of multidimensional 
-functions demonstrate the efficiency of the suggested approach.
+functions demonstrate the efficiency of the approach proposed by the authors.
 
@@
 
-The assumption of the objective function to be Lipschitzian is typical for many approaches to the development of the global optimization algorithms \cite{Evtushenko2013,Zilinskas2010,Pinter1996,Strongin2000}.
+The assumption of the objective function to be Lipschitzian is typical of many approaches to the development of the global optimization algorithms \cite{Evtushenko2013,Zilinskas2010,Pinter1996,Strongin2000}.
 Moreover, the adaptive estimate of the unknown Lipschitz constant, based on the obtained search information, is one of the most important problems being solved in these algorithms.
-The value of the Lipschitz constant affects the convergence rate of the global optimization algorithms essentially. Therefore, the issue of its correct estimate is so important. 
+The value of the Lipschitz constant affects essentially the convergence rate of the global optimization algorithms. Therefore, the issue of its correct estimate is so important. 
 The underestimation of the real value of this constant may result in losing the convergence of the algorithm to the global solution.
-At the same time, a too large value of the estimate of the constant could bring the objective function not matching its real behavior.
-This produces slow convergence of the algorithm to the global minimizer.
+At the same time, if the value of the constant estimate is too large, the objective function may not match its real behavior.
+This results in  slow convergence of the algorithm to the global minimizer.
 
@@
 
-Each approach from the above has its own advantages and disadvantages.
+Each of the above approaches has its own advantages and disadvantages.
 For example, the use of the global estimate over the whole search domain can slow down the convergence of the algorithm to the global minimizer.
-The use of the local estimates to accelerate the convergence of the method requires an adequate adjustment of the parameters of the algorithm in order to preserve the global convergence.
+The use of the local estimates to accelerate the convergence of the method requires an adequate adjustment of the algorithm parameters in order to preserve the global convergence.
 
-In the present work, it is considered a new algorithm in which there is the use of two global estimates of the Lipschitz constant. One of the two estimates is much greater than the other one. 
+In the present work, we  consider a new algorithm that uses two global estimates of the Lipschitz constant. One of the two estimates is much greater than the other one. 
 The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
@@
 
-The rigorous substantiation of the proposed approach goes beyond the present initial publication and will be done in the forthcoming works.
+A rigorous substantiation of the proposed approach goes beyond the present initial publication and will be done in the forthcoming works.
 Here we present the results of numerical experiments that clearly demonstrate the efficiency of the new algorithm.
-Several hundred multiextremal test problems of various dimensionalities have been solved conducting numerical experiments.
+Several hundred multiextremal test problems of various dimensionalities have been solved in the course of numerical experiments.
 
@@
 
-In the present paper, the approach based on the ideas of reduction of dimension with the use of the Peano-Hilbert curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously maps the unit interval $[0,1]$ onto the $N$-dimensional cube $D$ from (\ref{D}) has been used.
+In this paper, we follow the approach based on the idea of reducing the dimension with the use of the Peano-Hilbert curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously map the unit interval $[0,1]$ onto the $N$-dimensional cube $D$ from (\ref{D}) has been used.
 By using this kind of mapping, it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem
@@
 %Russian
-Отметим, что теоретически Peano-Hilbert curve $y(x)$ определяется as a limit object.
-Поэтому при практической реализации может быть построено лишь некоторое приближение к истинной space-filling curve.
-Методы построения подобных approximations (называемых evolvents) рассмотрены в \cite{Sergeyev2013,Strongin2000}.
+Note that theoretically the Peano-Hilbert curve $y(x)$ определяется as a limit object.
+is defined as a limit object.
+Therefore, in practical implementation, only some approximation to the true space-filling curve can be constructed.
+Some methods for constructing this type of approximations (called evolvents) are considered in \cite{Sergeyev2013,Strongin2000}.
 При этом точность приближения развертки к истинной кривой $y(x)$ зависит от плотности развертки $m$ (являющейся параметром ее построения) и составляет величину порядка $2^{-m}$ по каждой координате.
@@
 
-The global search algorithm presented in previous section is intended for solving the multiextremal problems, in which the objective function satisfies the Lipschitz condition.
+The global search algorithm presented in the previous section is intended for solving the multiextremal problems, in which the objective function satisfies the Lipschitz condition.
 It is not necessary to define the value of the constant for the algorithm convergence.
-The estimation of the constant is performed in the course of global search on the base of available search information.
+The estimation of the constant is performed in the course of global search based on available search information.
 According to the theorem from \cite{Strongin2000}, the sequence of the trial points $\{y^k\}$ will converge to the global minimizer $y^*$ if the condition 
@@
 
-Satisfying the condition (\ref{cond}) will be guaranteed at choosing large enough value of the parameter $r$.
+Satisfying the condition (\ref{cond}) will be guaranteed if we choose a large enough value of the parameter $r$.
 However, in this case the method will perform a large number of trials until the stop condition is satisfied.
@@
 An approach, in which two estimates of the Lipschitz constant %, one of which is essentially greater than another, 
-are used in the rules of the algorithm, is a promising one. 
-This approach implies the use of two parameters $r_{glob}$ and $r_{loc}$ where $r_{glob} > r_{loc}>1$ in the algorithm.
+are used in the rules of the algorithm, seems quite promising. 
+This approach implies the use in the algorithm of two parameters $r_{glob}$ and $r_{loc}$ where $r_{glob} > r_{loc}>1$.
 
 %Russian
-The rules of the algorithm with two estimates of the Lipschitz constant reproduce the ones of global search algorithm completely except Rule 3 (the computation of the characteristic) and Rule 4 (поиск интервала с максимальной характеристикой).
+The rules of the algorithm with two estimates of the Lipschitz constant reproduce the ones of the global search algorithm completely except Rule 3 (the computation of the characteristic) and Rule 4 (search for the interval with the maximum characteristic).
 
-New rule of calculating the characteristic $R(i)$ of the interval $(x_{i-1}, x_i)$ will consist of the following operations:
+The new rule for calculating the characteristic $R(i)$ of the interval $(x_{i-1}, x_i)$ will consist of the following operations:
 \begin{itemize}
@@
 %Russian
-Новые правила поиска интервала с максимальной характеристикой будут состоять в следующем:
+The new rules for finding the interval with the maximum characteristic will be as follows:
 \begin{itemize}
@@
 \item
-This value of $r$ is used in Rule 5 of the algorithm in the computing of the next trial point.
+Use this value of $r$ in Rule 5 of the algorithm in the computing of the next trial point.
 \end{itemize}
 
-This method of computing the characteristic is substantiated as follows. 
+This method for computing the characteristic is substantiated as follows. 
 %Russian
-На каждой итерации будет возникать интервал, на одной из границ которого достигается текущее минимальное значение $z^*$ from (\ref{z}).
-Причем на заключительной фазе поиска данный интервал будет соответствовать интервалу, содержащему глобальный минимум, т.е.
-будет являться лучшим с точки зрения проведения дальнейших испытаний в нем.
+Each search iteration will yield an interval with the current minimum value $z^*$ from (\ref{z}) at one of its boundaries.
+In the final phase of the search, this interval will correspond to the interval containing the global minimum, i.e.
+it will be the best one in terms of conducting further trials within it.
 
-Let current minimum value of $z^*$ from (\ref{z}) be achieved at the left point of the $i^{\rm th}$ interval, i.e.
-$z^* = z_{i-1}$.
-Then, according to the rule (\ref{R}) the following inequality will be true:
+Let the current minimum value of $z^*$ from (\ref{z}) be achieved at the left point of the $i^{\rm th}$ interval, i.e. $z^* = z_{i-1}$. 
+As proven in \cite{Strongin2000}, according to the rule (\ref{R}) the following inequality will be true: 
 \[
@@
 \]
-Следовательно, для нижних оценок характеристик $R_{loc}(i)$ and $R_{glob}(i)$ calculated with different parameters $r_{loc}$ and $r_{glob}$, будет выполняться соотношение
+Therefore, for evaluating the characteristics $R_{loc}(i)$ and $R_{glob}(i)$ calculated with different parameters $r_{loc}$ and $r_{glob}$, the following relation will hold:
 \[
@@
 
-Таким образом, если выбирать наибольую из характеристик в соответствии с
+Thus, when choosing the largest of the characteristics in accordance with
 \[
@@
 \]
-то, как правило, будет выбираться характеристика $R_{glob}(i)$, соответствующая большей оценке Lipschitz constant; меньшая оценка (ускоряющая процесс уточнения текущего решения) использоваться не будет.
-Однако если домножить нижнюю оценку для характеристики $R_{loc}(i)$ на коэффициент $\rho$ в соответствии с (\ref{pho}), то указанные нижние оценки будут равны, что сделает более вероятным выбор характеристики $R_{loc}(i)$, соответствующей меньшей оценке Lipschitz constant.
+%as a rule,
+the characteristic $R_{glob}(i)$ corresponding to the higher estimate of the Lipschitz constant will be chosen; the lower estimate (which speeds up the process of refining the current solution) will not be used.
+However, if we multiply the lower estimate for the characteristic $R_{loc}(i)$ by the coefficient $\rho$ in accordance with (\ref{pho}),then such lower estimates will be equal, 
+thus the choice of the characteristic $R_{loc}(i)$ corresponding to the lower estimate of the Lipschitz constant will become more likely.
 
@@
 %Russian
-Численное сравление алгоритмов было выполнено посредством решения нескольких серий задач, порожденных генератором GKLS \cite{Gaviano2003}.
-Данный генератор of multiextremal functions широко используется для сравнения методов глобальной оптимизации (see, for example, \cite{Barkalov2018,Paulavicius2014,Sergeyev2015}).
-В данном исследовании были решены шесть серий по 100 задач размерностей $N = 3,4,5$.
-Для каждой размерности генерировались \textit{Simple} and \textit{Hard} problems, отличающиеся размерами attraction region локальных и глобального экстремума.
-Задача считалась решенной, если метод проводил очередное испытание в точке $y^k$, которая находилась в $\delta$-октрестности global minimizer $y^\ast$, 
-т.е. $\left\|y^k-y^\ast\right\| <\delta\left\|b-a\right\|$, where  $a$ and $b$ --- границы search domain $D$.
+A numerical comparison of the algorithms was carried out by solving several series of problems generated by the GKLS generator \cite{Gaviano2003}.
+This generator of multiextremal functions is widely used to compare global optimization methods (see, for example, \cite{Barkalov2018,Paulavicius2014,Sergeyev2015}).
+In this study, six series each containing 100 problems of dimensions $N = 3,4,5$ were solved.
+For each dimension, \textit{Simple} and \textit{Hard} problems were generated, differing in the size of the attraction region for local extremums and global extremum.
+The problem was considered solved if the method conducted the next trial in sequence at a point that was in the $\delta$-neighborhood of the global minimizer $y^\ast$, 
+i.e. $\left\|y^k-y^\ast\right\| <\delta\left\|b-a\right\|$, where  $a$ and $b$ are the boundaries of the search domain $D$.
 
@@
 The averaged numbers of iterations performed by the algorithms are presented in Table \ref{tab:1}.
-At that, for GSA the values of parameter $r=4.8$ when solving the problems of Simple class and $r=5.6$ when solving the problems of Hard class were used. 
+For the GSA method the values of the parameter $r=4.8$ when solving the problems of the \textit{Simple} class and $r=5.6$ when solving the problems of the \textit{Hard} class were used. 
 These values are the minimum ones (with the accuracy 0.1), at which all problems have been solved successfully.
-When solving the problems from the above classes by GSA-DL, the value of parameter $r$ specified above was selected for the upper estimate of the Lipschitz constant, i.e.
+When solving the problems from the above classes by the GSA-DL method, the value of the parameter $r$ specified above was selected for the upper estimate of the Lipschitz constant, i.e.
 the value $r_{glob} = r$ was set, which was complemented by the values $r_{loc}=1.8$, $r_{loc}=2.1$ and $r_{loc}=2.4$.
@@
 
-The advantage of the GSA-DL algorithm over its prototype is confirmed by the operational characteristics of the algorithms as well.
+The advantages of the GSA-DL algorithm over its prototype are also confirmed by the operational characteristics of the algorithms as well.
 
@@
 The operational characteristics for the GSA and GSA-DL methods obtained when solving the \textit{Simple} and \textit{Hard} problem series with the dimensionalities $N=4$ and $N=5$ are presented in Fig.
-\ref{oper4} and \ref{oper5}, respectively.
+\ref{oper4} and Fig.
+\ref{oper5}, respectively.
 The values of the parameters $r$ used for estimating the Lipschitz constant are given in the figures.
@@
 The lower curves in Fig.
-\ref{oper4} and \ref{oper5} feature GSA whereas the upper ones --- GSA-DL.
-Such positions of the curves show the algorithm with two estimates of the Lipschitz constant provides much faster solving of the problem series in average that the algorithm using single estimate of the constant.
+\ref{oper4} and Fig.
+\ref{oper5} feature the characteristics of GSA whereas the upper ones, those of the GSA-DL.
+Such relative positions of the curves show the algorithm with two estimates of the Lipschitz constant is much faster on average when solving the problem series than the algorithm using a single estimate of the constant.
 %Russian
-Отметим, что для решения задач всех классов (кроме класса  \textit{Hard} при $N=5$) метод GSA-DL требует примерно в два раза меньше испытаний, чем GSA.
-Ухудшение результатов на классе \textit{Hard} при $N=5$ объясняется сложностью функций данного класса, характеризующихся большими attraction region локальных минимумов и узкой attraction region глобального минимума.
-Для корректного решения таких задач метод  GSA-DL чаще использует бОльшую оценку константы Липшица, что сближает скорость работы алгоритмов GSA и GSA-DL.
+Note that to solve problems of all classes (except for the \textit{Hard} class at $N=5$) the GSA-DL method requires about half as many trials as the GSA.
+The deterioration of the results in the case of the \textit{Hard} class at $N=5$ is explained by the complexity of this class' functions, which have a large attraction region of local minima and a narrow attraction region of the global minimum.
+To correctly solve such problems, the GSA-DL method often uses a higher estimate of the Lipschitz constant, thus reducing the speed difference of the GSA and GSA-DL algorithms.
 

### 2019-09-10 2019/NUMTA_Barkalov/LNCS/numta19.tex
Semi-final

@@
 The underestimation of the real value of this constant may result in losing the convergence of the algorithm to the global solution.
-At the same time, if the value of the constant estimate is too large, the objective function may not match its real behavior.
-This results in  slow convergence of the algorithm to the global minimizer.
+At the same time, if the value of the constant estimate for the objective function is too large and does not match its real behavior, this will slow down the convergence of the algorithm to the global minimizer.
 
@@
 
-In this paper, we follow the approach based on the idea of reducing the dimension with the use of the Peano-Hilbert curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously map the unit interval $[0,1]$ onto the $N$-dimensional cube $D$ from (\ref{D}) has been used.
+In this paper, we follow the approach based on the idea of reducing the dimension with the use of the Peano-Hilbert curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously map the unit interval $[0,1]$ onto the $N$-dimensional cube $D$ from (\ref{D}).
 By using this kind of mapping, it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem
@@
 \[
-\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N},
+\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}
 \]
@@
 %Russian
-Note that theoretically the Peano-Hilbert curve $y(x)$ определяется as a limit object.
-is defined as a limit object.
+Note that theoretically the Peano-Hilbert curve $y(x)$ is defined as a limit object.
 Therefore, in practical implementation, only some approximation to the true space-filling curve can be constructed.
 Some methods for constructing this type of approximations (called evolvents) are considered in \cite{Sergeyev2013,Strongin2000}.
-При этом точность приближения развертки к истинной кривой $y(x)$ зависит от плотности развертки $m$ (являющейся параметром ее построения) и составляет величину порядка $2^{-m}$ по каждой координате.
+In this case, the accuracy of the evolvent approximation to the true curve $y(x)$ depends on the density of the evolvent $m$ (which is a parameter for constructing the evolvent) and is of the order of $2^{-m}$ for each coordinate.
 
@@
 \begin{equation}\label{cond}
-r\mu > 2^{3-1/N}L\sqrt{N+3}.
+r\mu > 2^{3-1/N}L\sqrt{N+3}
 \end{equation}
@@
 are used in the rules of the algorithm, seems quite promising. 
-This approach implies the use in the algorithm of two parameters $r_{glob}$ and $r_{loc}$ where $r_{glob} > r_{loc}>1$.
+This approach implies the use in the algorithm of two parameters $r_{glob}$ and $r_{loc}$, where $r_{glob} > r_{loc}>1$.
 
@@
 
-This method for computing the characteristic is substantiated as follows. 
+This method for computing the characteristic can be substantiated as follows. 
 %Russian
@@
 \]
-Therefore, for evaluating the characteristics $R_{loc}(i)$ and $R_{glob}(i)$ calculated with different parameters $r_{loc}$ and $r_{glob}$, the following relation will hold:
+Therefore, for the estimates of the characteristics $R_{loc}(i)$ and $R_{glob}(i)$ calculated with different parameters $r_{loc}$ and $r_{glob}$, the following relation will hold:
 \[
@@
 the characteristic $R_{glob}(i)$ corresponding to the higher estimate of the Lipschitz constant will be chosen; the lower estimate (which speeds up the process of refining the current solution) will not be used.
-However, if we multiply the lower estimate for the characteristic $R_{loc}(i)$ by the coefficient $\rho$ in accordance with (\ref{pho}),then such lower estimates will be equal, 
+However, if we multiply the lower estimate for the characteristic $R_{loc}(i)$ by the coefficient $\rho$ in accordance with (\ref{pho}), then such lower estimates will be equal, 
 thus the choice of the characteristic $R_{loc}(i)$ corresponding to the lower estimate of the Lipschitz constant will become more likely.
@@
 The advantages of the GSA-DL algorithm over its prototype are also confirmed by the operational characteristics of the algorithms as well.
-
 Assume a series of test problems to be solved.
@@
 \ref{oper4} and Fig.
-\ref{oper5} feature the characteristics of GSA whereas the upper ones, those of the GSA-DL.
+\ref{oper5} feature the characteristics of the GSA method whereas the upper ones, those of the GSA-DL.
 Such relative positions of the curves show the algorithm with two estimates of the Lipschitz constant is much faster on average when solving the problem series than the algorithm using a single estimate of the constant.

### 2019-09-11 2019/NUMTA_Barkalov/LNCS/numta19.tex
Italic font for Simple and Hard

@@
 \end{minipage}
-\caption{Operational characteristics for GKLS Simple (a) and Hard (b) classes, $N=4$.}
+\caption{Operational characteristics for GKLS \textit{Simple} (a) and \textit{Hard} (b) classes, $N=4$.}
 \label{oper4}
@@
 \end{minipage}
-\caption{Operational characteristics for GKLS Simple (a) and Hard (b) classes, $N=5$.}
+\caption{Operational characteristics for GKLS \textit{Simple} (a) and \textit{Hard} (b) classes, $N=5$.}
 \label{oper5}

### 2019-09-11 2019/NUMTA_Barkalov/LNCS/numta19.tex
Four extra lines

@@
 	\item local estimations of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ \cite{Kvasov2003,Sergeyev2010,Sergeyev2016}.
-	\item the choice of the estimates of the constant $L$ from a set of possible values \cite{Gablonsky2001,Jones1993,Jones2009,Sergeyev2006}.
+	\item the choice of the estimates of the constant $L$ from a set of possible values \cite{Gablonsky2001,Jones1993,Sergeyev2006}.
 \end{itemize}
@@
 \item 
-Select the interval $(x_{t-1},x_t)$ corresponding to the maximum characteristic
-\[
-R(t)= \max_{1 \leq i \leq k}R(i).
-\]
+Select the interval $(x_{t-1},x_t)$ corresponding to the maximum characteristic $R(t)= \max_{1 \leq i \leq k}R(i)$.
 \item
@@
 \[
-R(i) \geq \Delta_i \left( 1 - \frac{1}{r} \right)^2.
+R(i) \geq \Delta_i \left( 1 - 1/r \right)^2.
 \]
@@
 \[
-	\Delta_i \left( 1 - \frac{1}{r_{glob}} \right)^2 > \Delta_i \left( 1 - \frac{1}{r_{loc}} \right)^2.
+	\Delta_i \left( 1 - 1/r_{glob} \right)^2 > \Delta_i \left( 1 - 1/r_{loc} \right)^2.
 \]
@@
 
+The advantages of the GSA-DL algorithm over its prototype are also confirmed by the operational characteristics of the algorithms as well.
+Assume a series of test problems to be solved.
+The results of solving the series can be presented by a function $p(k)$ featuring the fraction of the total number of problems solved in $k$ iterations.
+Such a function will be called the \textit{operational characteristic} of the algorithm.
+
+The operational characteristics for the GSA and GSA-DL methods obtained when solving the \textit{Simple} and \textit{Hard} problem series with the dimensionalities $N=4$ and $N=5$ are presented in Fig.
+\ref{oper4} and Fig.
+\ref{oper5}, respectively.
+The values of the parameters $r$ used for estimating the Lipschitz constant are given in the figures.
+
 \begin{figure}
@@
 \end{figure}
-
-The advantages of the GSA-DL algorithm over its prototype are also confirmed by the operational characteristics of the algorithms as well.
-Assume a series of test problems to be solved.
-The results of solving the series can be presented by a function $p(k)$ featuring the fraction of the total number of problems solved in $k$ iterations.
-Such a function will be called the \textit{operational characteristic} of the algorithm.
-
-The operational characteristics for the GSA and GSA-DL methods obtained when solving the \textit{Simple} and \textit{Hard} problem series with the dimensionalities $N=4$ and $N=5$ are presented in Fig.
-\ref{oper4} and Fig.
-\ref{oper5}, respectively.
-The values of the parameters $r$ used for estimating the Lipschitz constant are given in the figures.
 

### 2019-09-12 2019/NUMTA_Barkalov/LNCS/numta19.tex
8 pages and vector figures

@@
 A numerical comparison of the algorithms was carried out by solving several series of problems generated by the GKLS generator \cite{Gaviano2003}.
-This generator of multiextremal functions is widely used to compare global optimization methods (see, for example, \cite{Barkalov2018,Paulavicius2014,Sergeyev2015}).
+This generator of multiextremal functions is widely used to compare global optimization methods (see, for example, \cite{Barkalov2018,Sergeyev2006,Sergeyev2015}).
 In this study, six series each containing 100 problems of dimensions $N = 3,4,5$ were solved.
-For each dimension, \textit{Simple} and \textit{Hard} problems were generated, differing in the size of the attraction region for local extremums and global extremum.
-The problem was considered solved if the method conducted the next trial in sequence at a point that was in the $\delta$-neighborhood of the global minimizer $y^\ast$, 
+For each dimension, \textit{Simple} and \textit{Hard} problems were generated, differing in the size of the attraction regions for local extremums and global extremum.
+The problem was considered solved if the method conducted the trial at a point that was in the $\delta$-neighborhood of the global minimizer $y^\ast$, 
 i.e. $\left\|y^k-y^\ast\right\| <\delta\left\|b-a\right\|$, where  $a$ and $b$ are the boundaries of the search domain $D$.
@@
 \begin{figure}
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklss_4.png} \\ (a)}
+\begin{minipage}{0.47\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklss_4.pdf} {\scriptsize \\ (a)}}
 \end{minipage}
 \hfill
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklsh_4.png} \\ (b)}
+\begin{minipage}{0.47\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklsh_4.pdf} {\scriptsize \\ (b)}}
 \end{minipage}
@@
 \begin{figure}
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklss_5.png} \\ (a)}
+\begin{minipage}{0.47\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklss_5.pdf} {\scriptsize \\ (a)}}
 \end{minipage}
 \hfill
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklsh_5.png} \\ (b)}
+\begin{minipage}{0.47\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{Operating_characteristic_gklsh_5.pdf} {\scriptsize \\ (b)}}
 \end{minipage}

### 2019-09-12 2019/NUMTA_Barkalov/LNCS/numta19.tex
Reliabiliry

@@
 \end{equation} 
-and the real number $r>1$ is 
-the input parameter of the algorithm.
+and the real number $r>1$ is a reliability parameter of the algorithm.
 \item 
@@
 
-An approach, in which two estimates of the Lipschitz constant %, one of which is essentially greater than another, 
-are used in the rules of the algorithm, seems quite promising. 
+An approach, in which two estimates of the Lipschitz constant are used in the rules of the algorithm, seems quite promising. 
 This approach implies the use in the algorithm of two parameters $r_{glob}$ and $r_{loc}$, where $r_{glob} > r_{loc}>1$.

### 2019-09-25 2019/Automation/Barkalov.tex
Fix some misprints

@@
 Способ решения задачи (\ref{main_problem}) на основе соотношения (\ref{block_nested}) будет в целом повторять  (\ref{nested_N})--(\ref{nested_1}).
-Необходимо лишь заменить исходные переменны $y_i, 1\leq i \leq N$,  на блочные переменные variables $u_i, 1\leq i \leq M$.
+Необходимо лишь заменить исходные переменны $y_i, 1\leq i \leq N$,  на блочные переменные $u_i, 1\leq i \leq M$.
 
@@
 Тогда каждой подзадаче (\ref{block_nested_i}) можно присвоить некоторое числовое значение, называемое характеристикой этой задачи.
-К качестве такой характеристики можно взять значение $R(t)$ из (\ref{MaxR}), т.е.
+В качестве такой характеристики можно взять значение $R(t)$ из (\ref{MaxR}), т.е.
 максимальную характеристику интервалов, сформированных в данной задаче.
@@
 На рис.~\ref{fig2}(a,b) приведены операционные характеристики алгоритмов, полученные на указанных классах задач.
-Непрерывная линия соответствует алгоритму с использованием разверток, короткий пунктир --- адаптивной многошаговой схемы, длинный пунктир --- многошаговой схемы.
+Непрерывная линия соответствует алгоритму с использованием разверток, короткий пунктир --- адаптивной многошаговой схеме, длинный пунктир --- многошаговой схеме.
 Результаты экспериментов показывают, что АГП с использованием адаптивной многошаговой схемы показывает примерно одинаковое быстродействие  по сравнению с развертками, и оба они значительно превосходят алгоритм, использующий многошаговую схему.

### 2019-09-26 2019/Automation/Barkalov.tex
Fix some misprints

@@
 В статье рассматриваются задачи многомерной многоэкстремальной оптимизации и численные методы их решения.
-Об оптимизируемой функции делается лишь общее предположение, что она удовлетворяет условию Липшица с априори неизвестной константой.
-Задачи такого типа часто встречаются в приложениях.
+Об оптимизируемой функции делается лишь общее предположение, что она удовлетворяет условию Липшица с априори неизвестной константой (задачи такого типа часто встречаются в приложениях).
 Рассмотрено два способа редукции размерности в задачах многомерной оптимизации: ипользование кривых Пеано (разверток) и рекурсиваная многошаговая схема.
@@
 Например, многие известные методы основаны на различных способах разбиения области поиска на систему подобластей и последующего выбора наиболее перспективной подобласти для размещения очередного испытания (вычисления целевой функции).
-Результаты в данном направлении представлены в работах \cite{Jones09,Zilinskas10,Evtushenko13,Kvasov13,Paulavicius16}.
+Результаты, полученные в данном направлении, представлены в работах \cite{Jones09,Zilinskas10,Evtushenko13,Kvasov13,Paulavicius16}.
 
@@
 Разработанные методы основаны на редукции исходной многомерной задачи к эквивалентной одномерной или системе одномерных подзадач и последующим решением одномерных задач эффективными методами оптимизации функций одной переменной.
-Предложено две такие схемы: редукция на основе кривых, заполняющих пространство (кривых Пеано, или разверток) \cite{Strongin91,Strongin00}, и схема рекурсивной вложенной оптимизации (многошаговая схема) \cite{Grishagin97,Grishagin01}.
+Предложено две такие схемы: редукция на основе кривых, заполняющих пространство (кривых Пеано, или \textit{разверток}) \cite{Strongin91,Strongin00}, и схема рекурсивной вложенной оптимизации (\textit{многошаговая схема}) \cite{Grishagin97,Grishagin01}.
 В статье \cite{Grishagin16} предложена адаптивная многошаговая схема, существенно повышающая эффективность оптимизации по сравнению с базовым прототипом.
 В данной работе предлагается обобщение адаптивной схемы редукции размерности, комбинирующее использование вложенной оптимизации и кривых Пеано.
-При таком подходе вложенные подзадачи в адаптивной схеме могут быть как одномерными, так и многомерными.
-В последнем случае для редукции размерности вложенных подзадач используются развертки.
+При таком подходе вложенные подзадачи в адаптивной схеме могут быть как одномерными, так и многомерными; в последнем случае для редукции размерности вложенных подзадач используются развертки.
 
@@
 \end{equation}
-с целевой функцией, удовлетворяющей условию Липшица
+с целевой функцией, удовлетворяющей условию Липшица.
 
-Приведем описание алгоритма глобального поиска (АГП) в соответствии с \cite{Strongin00}.
+Приведем описание алгоритма глобального поиска (АГП) для решения базовой задачи в соответствии с \cite{Strongin00}.
 В процессе своей работы АГП породждает последовательность точек $x^i$, в которых вычисляются значения целевой функции $z^i=\varphi(x^i)$. 
@@
 
-В соответствии с алгоритмом, первые два испытания проводятся в граничных точках интервала $[a,b]$, т.е., $x^0=a,\;x^1=b$. 
+В соответствии с алгоритмом, первые два испытания проводятся в граничных точках интервала $[0,1]$, т.е. $x^0=0,\;x^1=1$. 
 В этих точка вычисляются значения целевой функции $z^0=\varphi(x^0),\;z^1=\varphi(x^1)$, и устанавливается значение счетчика $k=1$. 
@@
 R(t)=\max_{1\leq i\leq {k}}R(i).
-\end{equation}.
+\end{equation}
 Если максимальная характеристика соответствует нескольким интервалам, то в качестве $t$ выбрать минимальное число, удовлетворяющее (\ref{MaxR}).
@@
 \begin{equation}
-\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq N\right\}=\left\{
-y(x):0\leq x \leq 1 \right\}
+\left\{y(x):0\leq x \leq 1 \right\} = \left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq N\right\}.
 \end{equation}
 
-Отметим, что теоретически кривая $y(x)$ определяется как предельнгый объект.
+Отметим, что теоретическая кривая $y(x)$ определяется как предельный объект.
 Поэтому при практической реализации может быть построено лишь некоторое приближение к истинной кривой.
-Методы построения подобных приближений (называемых \textit{развертками}) рассмотрены в \cite{Strongin78,Strongin00}.
-При этом точность приближения развертки к истинной кривой $y(x)$ зависит от \textit{плотности развертки} $m$ (являющейся параметром ее построения) и составляет величину порядка $2^{-m}$ по каждой координате.
+Методы построения подобных аппроксимаций (называемых \textit{развертками}) рассмотрены в \cite{Strongin78,Strongin00}.
+При этом точность приближения развертки к истинной кривой $y(x)$ зависит от \textit{плотности} развертки $m$ (являющейся параметром ее построения) и составляет величину порядка $2^{-m}$ по каждой координате.
 
-Используя подобного род отображения можно свести решение многомерной задачи (\ref{main_problem}) к решению эквивалетной ей одномерной
+Использование подобного рода отображений позволяет свести решение многомерной задачи (\ref{main_problem}) к решению эквивалентной ей одномерной
 \begin{equation}
@@
 Условия (\ref{Holder}), (\ref{HL}) позволяет легко обобщить <<одномерный>> алгорим из раздела \ref{SectionCore} для решения многомерных задач. 
-Для этого достаточно заменить длины интервалов $\Delta_i$ в формулах (\ref{mu}),(\ref{R}) на длины, вычисляемые в соответствии с 
+Для этого достаточно заменить длины интервалов $\Delta_i$ в формулах (\ref{mu}),(\ref{R}) на длины 
 \begin{equation}
@@
 Задачу будем считать решенной, если алгоритм сгенерировал точку $y^k$ очередного испытания в окрестности решения $y^\ast$, т.е.
-$\left|y^k-y^\ast\right| < \delta \left\|b-a\right\|$, где $\delta = 0.01$, $a$ и $b$ --- границы области поиска $D$.
+$\left\|y^k-y^\ast\right\| < \delta \left\|b-a\right\|$, где $\delta = 0.01$, $a$ и $b$ --- границы области поиска $D$.
 
@@
 Непрерывная линия соответствует алгоритму с использованием разверток, короткий пунктир --- адаптивной многошаговой схеме, длинный пунктир --- многошаговой схеме.
-Результаты экспериментов показывают, что АГП с использованием адаптивной многошаговой схемы показывает примерно одинаковое быстродействие  по сравнению с развертками, и оба они значительно превосходят алгоритм, использующий многошаговую схему.
+Результаты экспериментов показывают, что АГП с использованием адаптивной многошаговой схемы демонстрирует примерно одинаковое быстродействие  по сравнению с развертками, и оба они значительно превосходят алгоритм, использующий многошаговую схему.
 Поэтому в дальнейших экспериментах мы ограничимся сравнением различных вариантов адаптивной схемы редукции размерности.
@@
 Сравнивалась работа АГП с использованием разверток и блочной адаптивной схемы с формированием двух уровней подзадач одинаковой размерности $N_1=N_2=3$.
-Среднее число испытаний, выполненных АГП с использованием разверток, составило 102987, тогда как с использованием блочной адаптивной схемы  --- 75 390. 
+Среднее число испытаний, выполненных АГП с использованием разверток, составило 102 987, тогда как с использованием блочной адаптивной схемы  --- 75 390. 
 На рис.~\ref{fig4} приведены операционные характеристики алгоритмов. Пунктирная линия соответствует алгоритму с использованием разверток, а непрерывная --- блочной адаптивной схемы.
@@
 В данной работе предложена обобщенная адаптивная схема редукции размерности для задач глобальной оптимизации, комбинирующая использование кривых Пеано и схему вложенной (рекурсивной) оптимизации.
-Для решения редуцированных подзадач меньшей размерности используется алгоритм глобального поиска.
+Для решения редуцированных подзадач используется алгоритм глобального поиска.
 Приведена вычислительная схема алгоритма, рассмотрены основные вопросы, связанные с использованием адаптивной схемы редукции размерности.
@@
 \bibitem{Jones09}%Проверить формат ссылки!!!
-{\it Jones D. R.} The direct global optimization algorithm. In: Floudas C. A., Pardalos P. M. (eds.) The Encyclopedia of Optimization, Second Edition.  Springer. 2009. P. 725--735.
+{\it Jones D. R.} The DIRECT global optimization algorithm. In: Floudas C. A., Pardalos P. M. (eds.) The Encyclopedia of Optimization, Second Edition.  Springer. 2009. P. 725--735.
 
@@
 \bibitem{Grishagin2001}
-{\it Sergeyev Y., Grishagin V.} Parallel asynchronous global search and the nested
-optimization scheme // J. Comput. Anal. Appl. 2001. V. 3(2). P. 123--145.
+{\it Sergeyev Y., Grishagin V.} Parallel asynchronous global search and the nested optimization scheme // J. Comput. Anal. Appl. 2001. V. 3(2). P. 123--145.
 
 \bibitem{Grishagin2015}
-{\it Gergel V., Grishagin V., Israfilov R.} Local tuning in nested scheme of global
-optimization // Procedia Computer Science. 2015. V. 51(1). P. 865--874
+{\it Gergel V., Grishagin V., Israfilov R.} Local tuning in nested scheme of global optimization // Procedia Computer Science. 2015. V. 51(1). P. 865--874
 
 \bibitem{Barkalov2016}
-{\it Barkalov K., Lebedev I.} Solving multidimensional global optimization problems
-using graphics accelerators // CCIS. 2016. V. 687. P. 224--235.
+{\it Barkalov K., Lebedev I.} Solving multidimensional global optimization problems using graphics accelerators // CCIS. 2016. V. 687. P. 224--235.
 

### 2019-09-26 2019/Automation/Barkalov.tex
Fix some misprints

@@
 Об оптимизируемой функции делается лишь общее предположение, что она удовлетворяет условию Липшица с априори неизвестной константой (задачи такого типа часто встречаются в приложениях).
-Рассмотрено два способа редукции размерности в задачах многомерной оптимизации: ипользование кривых Пеано (разверток) и рекурсиваная многошаговая схема.
+Рассмотрено два способа редукции размерности в задачах многомерной оптимизации: использованные кривых Пеано (разверток) и рекурсивная многошаговая схема.
 Предложена обобщенная схема, комбинирующая эти два подхода.
@@
 
-Многошаговая схема редукции размерности (схема вложненной оптимизации) основана на известном соотношении %ссылка!!!! Carr? 
+Многошаговая схема редукции размерности (схема вложенной оптимизации) основана на известном соотношении %ссылка!!!! Carr? 
 \begin{equation}\label{nested}
@@
 
-Для описанной выше многошаговой схемы было предложено обобщение --- \textit{блочная многошаговая схема}, которое комбинирует использование разверток и вложенной оптмизации \cite{Barkalov2014}.
+Для описанной выше многошаговой схемы было предложено обобщение --- \textit{блочная многошаговая схема}, которое комбинирует использование разверток и вложенной оптимизации \cite{Barkalov2014}.
 
@@
 Способ решения задачи (\ref{main_problem}) на основе соотношения (\ref{block_nested}) будет в целом повторять  (\ref{nested_N})--(\ref{nested_1}).
-Необходимо лишь заменить исходные переменны $y_i, 1\leq i \leq N$,  на блочные переменные $u_i, 1\leq i \leq M$.
+Необходимо лишь заменить исходные переменные $y_i, 1\leq i \leq N$,  на блочные переменные $u_i, 1\leq i \leq M$.
 
@@
 Решение возникающего множества подзадач вида (\ref{nested_i}) (для многошаговой схемы) или вида (\ref{block_nested_i}) (для блочной многошаговой схемы) может быть организовано различными способами. 
-Очевидный способ (детально проработанный в \cite{Grishagin97,Grishagin2001,Grishagin2015} для многошаговой схемы и в \cite{Barkalov2014,Barkalov2016} для блочной многошаговой схемы) основан на решении поздадач в соответствии с рекурсивным порядком их порождения.
+Очевидный способ (детально проработанный в \cite{Grishagin97,Grishagin2001,Grishagin2015} для многошаговой схемы и в \cite{Barkalov2014,Barkalov2016} для блочной многошаговой схемы) основан на решении подзадач в соответствии с рекурсивным порядком их порождения.
 Однако здесь возникает потеря значительной части информации о целевой функции.
@@
 Иным  подходом является адаптивная схема, в которой все подзадачи решаются одновременно, что позволяет более полно учитывать информацию о многомерной задаче и за счет этого ускорять процесс ее решения.
-Для случая одномерных поздадач данный подход был теоретически обоснован и апробирован в \cite{Grishagin16,Grishagin2016_1,Grishagin18}.
+Для случая одномерных подзадач данный подход был теоретически обоснован и апробирован в \cite{Grishagin16,Grishagin2016_1,Grishagin18}.
 Настоящая работа предлагает обобщение адаптивной схемы для случая многомерных подзадач.

### 2019-09-27 2019/Automation/Barkalov.tex
Fix the title and references

@@
 \documentclass[12pt]{a&t}
-\usepackage{hyperref}
+%\usepackage{hyperref}
 \usepackage{graphicx}
@@
 \year{2019}
-\title{АДАПТИВНАЯ ГЛОБАЛЬНАЯ ОПТИМИЗАЦИИ НА ОСНОВЕ БЛОЧНО-РЕКУРСИВНОЙ СХЕМЫ РЕДУКЦИИ РАЗМЕРНОСТИ}%
+\title{АДАПТИВНАЯ ГЛОБАЛЬНАЯ ОПТИМИЗАЦИЯ НА ОСНОВЕ БЛОЧНО-РЕКУРСИВНОЙ СХЕМЫ РЕДУКЦИИ РАЗМЕРНОСТИ}%
 \thanks{Исследование выполнено за счет гранта Российского научного фонда (проект \mbox{№\,16-11-10150}).}
@@
 Приведем описание алгоритма глобального поиска (АГП) для решения базовой задачи в соответствии с \cite{Strongin00}.
-В процессе своей работы АГП породждает последовательность точек $x^i$, в которых вычисляются значения целевой функции $z^i=\varphi(x^i)$. 
+В процессе своей работы АГП порождает последовательность точек $x^i$, в которых вычисляются значения целевой функции $z^i=\varphi(x^i)$. 
 Будем называть процесс вычисления значения целевой функции \textit{испытанием}.
@@
 
-Условия (\ref{Holder}), (\ref{HL}) позволяет легко обобщить <<одномерный>> алгорим из раздела \ref{SectionCore} для решения многомерных задач. 
+Условия (\ref{Holder}), (\ref{HL}) позволяет легко обобщить <<одномерный>> алгоритм из раздела \ref{SectionCore} для решения многомерных задач. 
 Для этого достаточно заменить длины интервалов $\Delta_i$ в формулах (\ref{mu}),(\ref{R}) на длины 
@@
 
-Многошаговая схема редукции размерности (схема вложенной оптимизации) основана на известном соотношении %ссылка!!!! Carr? 
+Многошаговая схема редукции размерности (схема вложенной оптимизации) основана на известном соотношении \cite{Strongin78,Strongin13} 
 \begin{equation}\label{nested}
@@
 \end{equation}
-которое позволяет свести решение исходной многомерной задачи (\ref{main_problem}) к решению семейства рекурсивно связанных одномерных позадач.
+которое позволяет свести решение исходной многомерной задачи (\ref{main_problem}) к решению семейства рекурсивно связанных одномерных подзадач.
 

### 2019-09-28 2019/RSD_Poster/poster.tex
Add RSD Poster

--- 2017-03-08 2017/PAVT/poster/poster.tex
@@
 \usepackage{fontspec}
-\setmainfont{CMU Sans Serif}
-\setromanfont{CMU Sans Serif}
-\setsansfont{CMU Sans Serif}
+
+\setmainfont{Times New Roman}
+\setromanfont{Times New Roman}
+\setsansfont{Times New Roman}
 
@@
 \setmainlanguage{russian}
+\usepackage{csquotes}
 
@@
             autolang=other]{biblatex}
+
+%\newfontfamily\cyrillicfonttt{lmmonolt10-regular.otf}
+%\newfontfamily\cyrillicfontsf{lmsans10-regular.otf}
+
 \addbibresource{bibliography.bib}
@@
 
-\title{Использование параллельной системы глобальной оптимизации Globalizer для
-решения задач оптимального управления}
-\author{И.Г. Лебедев \and В. В. Соврасов}
+\title{Эффективные методы глобальной оптимизации
+для решения задач оптимального управления
+}
+\author{К.А. Баркалов \and М.А. Кочеганова \and С.А. Бевзюк \and А.А. Федюков}
 \institute{ННГУ им. Н.И. Лобачевского}
@@
         \begin{column}[t]{0.48\paperwidth}
-            \begin{block}{Задача поиска оптимального управления}
-              Необходимо найти оптимальное управление в виде обратной связи по состоянию для системы:
-              \begin{displaymath}
-                \dot x = (A+B_u\Theta)x + B_v v, x(0)=0
-              \end{displaymath}
-              Выходы системы описываются выражениями:
-              \begin{displaymath}
-                z_k=(C_k+B_u\Theta),k=\overline{1,N}
-              \end{displaymath}
-              Критерии оптимальности:
-              \begin{displaymath}
-                J_k(\Theta)=\sup_{v\in L_2} \frac{\max_{1\leqslant i \leqslant n_k} \sup_{t\geqslant 0}|z_k^{(i)}(\Theta,t)|}{||v||_2} \rightarrow\min_{\Theta}
-              \end{displaymath}
-            Ограничение на устойчивость системы:
-            \begin{displaymath}
-              g_0(\Theta)=\min_{j}\re(\lambda_j(A+B_u\Theta)) < 0
-            \end{displaymath}
-            Многокритериальная задача сводится к скалярной методом уступок или с помощью свёртки Гермейера.
-            В \cite{optControl} доказано, что использование последней позволяет найти всё множество Парето.
+            \begin{block}{Задача поиска $H_{\infty}$-динамического регулятора}
+              Рассматривается задача поиска $H_{\infty}$ -динамического регулятора для обратного маятника с подвижным основанием.
+							\begin{equation}\label{eq:system}
+                \dot{x}=A x + B_1 v + B_2 u
+              \end{equation}
+							Известно \cite{synthesisControlBook}, что для существования регулятора с наименьшим уровнем гашения возмущений $\gamma>0$ необходимо и достаточно, чтобы существовала матрица $X=X^T>0$, удовлетворяющая  двум матричным неравенствам, одно из которых не является выпуклым, т.к.
+зависит от обратной матрицы $X^{-1}$:
+							\begin{equation}\label{eq:matrix_1}
+                W_{P}^T 
+								\begin{bmatrix}
+                  \quad A_{0}^T X + X A_0 & X B_0 & C_{0}^T \\
+                  B_{0}^T X & -\gamma I & 0 \\
+									C_0  & 0 & -\gamma I \quad \\
+                \end{bmatrix}
+								W_P
+                < 0
+              \end{equation}
+							\begin{equation}\label{eq:matrix_2}
+                W_{R}^T 
+								\begin{bmatrix}
+                  \quad X^{-1} A_{0}^T + A_0 X^{-1} & B_0 & X^{-1} C_{0}^T \quad  \\
+                  B_{0}^T & -\gamma I & 0 \\
+									C_0 X^{-1} & 0 & -\gamma I\\
+                \end{bmatrix}
+								W_R
+                < 0
+              \end{equation}
+							Возникает задача поиска минимального значения $\gamma$  при ограничениях \eqref{eq:matrix_1} и \eqref{eq:matrix_2}.
+Причем неравенство \eqref{eq:matrix_2} не может быть представлено в виде линейного матричного неравенства, т.е. данную задачу нельзя решить стандартными методами выпуклой оптимизации в сочетании
+с методами решения линейных матричных неравенств, реализованными,
+например, в MATLAB Robust Toolbox. Требуется использование
+методов, обладающих сходимостью к глобальному экстремуму.
+
           \end{block}
             \begin{block}{Задача глобальной оптимизации}
-              Постановка задачи с ограничениями:
+              Общая постановка задачи глобальной оптимизации с ограничениями:
               \begin{displaymath}
@@
             \end{block}
-            \begin{block}{Метод глобальной оптимизмции}
-              Для оптимизации используется одномерный метод Стронгина.
-              Редукция размерности осуществляется с помощью кривой Пеано.
+            \begin{block}{Метод глобальной оптимизации}
+              Задача оптимизации решается Алгоритмом Глобального Поиска (АГП), предложенным Р.Г.
+Стронгиным \cite{strOptBook}.
+При этом решение многомерной задачи сводится к решению эквивалентной ей одномерной задачи с помощью развертки Пеано, отображающей отрезок $[0,1]$  на гиперкуб $D$.
 
               Общая схема одной итерации одномерного метода:
-              \begin{itemize}
+              \begin{enumerate}
                 \justifying
@@
                 \item Проверить выполнение критерия остановки \(x_{t}-x_{t-1}<\varepsilon\).
-              \end{itemize}
-              Подробное описание метода можно найти в \cite{optParallelBook}.
+              \end{enumerate}
+              Подробное описание АГП, а также специальной схемы учёта ограничений (без использования штрафных функций) можно найти в \cite{strOptBook}.
 
@@
               \end{minipage}
-
             \end{block}
@@
         \begin{column}[t]{0.48\paperwidth}
-          \begin{block}{Способ распараллеливания}
-
+          \begin{block}{Параллельные версии метода оптимизации}
+            Существует несколько способов распареллеливания алгоритма глобального поиска (АГП):
+            \begin{itemize}
+              \justifying
+              \item Распараллеливание по характеристикам в рамках системы с общей памятью.
+              На шаге 2 вместо одного интервала выбираются \(p\) интервалов с наилучшими характеристиками и в них параллельно проводятся испытания.
+              \item Распараллеливание по развёрткам в рамках системы с раздельной памятью. На каждом узле системы работает копия метода, использующая
+              уникальную развёртку. Копии метода обмениваются многомерными точками, однако одномерные прообразы этих точек различны для каждого метода.
+              При использовании \(L\) развёрток каждый метод дополнительно получает \(L-1\) точку в свою поисковую информацию на каждой итерации, что
+              ускоряет его сходимость.
+              \item Сочетание указанных выше подходов.
+            \end{itemize}
+          В \cite{optParallelBook} перечисленные схемы описаны более подробно.
           \end{block}
           \begin{block}{Результаты}
+
+						Решена модельная задача, связанная со стабилизацией плоского перевернутого маятника с подвижным основанием.
+Найдены параметры регулятора $\theta$ и соответствующее им минимальное значение уровня гашения возмущений $\gamma$. 
+Размерность (число параметров) возникающей при этом задачи глобальной оптимизации $N=7$.
+
+          \begin{table}[h]
+            \caption{Сравнение последовательного и параллельного алгоритма}
+            \label{tab:Results}
+            \begin{tabular}{l|c}
+						  \hline
+              Алгоритм & Время решения задачи (сек.) \\ 
+              \hline
+              Последовательный алгоритм  &  125 \\
+              Параллельный алгоритм (4 ядра)  &  82 \\
+              \hline
+							Ускорение & 1.5 \\
+							\hline
+            \end{tabular}
+           \end{table}
+
+					Эксперименты проводились в системе Globalizer \cite{globalizer} на узле кластера <<Лобачевский>>, CPU Intel Core i7 (3.6 GHz).
           \end{block}
-            \begin{block}{Ссылки}
+          \begin{block}{Дальнейшая работа}
+						\begin{enumerate}
+                \item В процессе решения задачи обнаружен эффект резкого роста константы Липшица для функций ограничений задачи, что существенно затрудняет поиск глобального экстремума.
+Планируется разработать подход к учету и обработке подобных ситуаций (идентификация и исключение из рассмотрения подобластей, в которых происходит скачкообразное изменение значений функции).
+
+                \item Вычисление ограничений задачи сводится к обращению матрицы $X$ и поиску собственных чисел матриц из правой части ограничений \eqref{eq:matrix_1}, \eqref{eq:matrix_2}.
+Планируется распараллелить данные операции, что будет актуальным при поиске динамических регуляторов для задач вида \eqref{eq:system} большой размерности.
+						\end{enumerate}
+          \end{block}
+          \begin{block}{Литература}
             \printbibliography

### 2019-10-01 2019/RSD_Integer/RSD2019.tex
RSD: add middots and letter symbol
Final
Add one reference for SuperFRI (Sulimov)

@@
-\documentclass{llncs}
+\documentclass[runningheads]{llncs}
 
@@
 
+\def\middot{\textperiodcentered~}
+
 \begin{document}
@@
 }
-\author{Konstantin Barkalov \and Ilya Lebedev %\Letter 
+\author{Konstantin Barkalov (\Letter) \and Ilya Lebedev %\Letter
 \\
@@
 
-\keywords { Global optimization, Non-convex constraints, Mixed-integer problems}
+\keywords { Global optimization \middot Non-convex constraints \middot Mixed-integer problems
+\middot Parallel algorithms}
 
@@
 general, a nonuniform one) of the space of parameters. 
+Such problems are common in applications \cite{Sulimov}.
 The problems, in which some parameters can take the discrete or integer values only (\textit{mixed-integer global
@@
 \begin{minipage}{0.49\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (a)}
+\center{\includegraphics[width=1.0\linewidth]{fig1b.jpg} \\ (a)}
 \end{minipage}
 \begin{minipage}{0.49\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (b)}
+\center{\includegraphics[width=1.0\linewidth]{fig1c.jpg} \\ (b)}
 \end{minipage}
@@
 
+\bibitem{Sulimov}
+Kutov, D.C., Sulimov, A.V., Sulimov, V.B.: Supercomputer docking: investigation of low energy minima of protein-ligand complexes.
+Supercomputing Frontiers and Innovations \textbf{5}(3), 134--137 (2018)
+
 \bibitem{Burer}

### 2019-10-01 2019/RSD_Integer/RSD2019.tex
Add two more references from SuperFRI

@@
 As a result, the search of the global optimum is reduced to the construction of a coverage (in
-general, a nonuniform one) of the space of parameters. 
-Such problems are common in applications \cite{Sulimov}.
+general, a nonuniform one) of the space of parameters; 
+such problems are common in applications \cite{Anisimov,Romanov,Sulimov}.
 The problems, in which some parameters can take the discrete or integer values only (\textit{mixed-integer global
@@
 
+\bibitem{Anisimov}
+Anisimov, K.S., Savelyev, A.A.,	Kursakov, I.A.,	Lysenkov, A.V.,	Prakasha, P.S.: Optimization of BWB aircraft using parallel computing.
+Supercomputing Frontiers and Innovations \textbf{5}(3), 93--97 (2018)
+
+\bibitem{Romanov}
+Romanov, S.Y.: Supercomputer simulations of nondestructive tomographic imaging with rotating transducers. Supercomputing Frontiers and Innovations \textbf{5}(3), 98--102 (2018)
+
 \bibitem{Sulimov}
@@
 \bibitem{Sovrasov2019}
-Sovrasov, V., Bevzuk, S.: Comparison of several sequential and parallel derivative-free global optimization algorithms. 	Communications in Computer and Information Science (2019) (article in press)
+Sovrasov, V.: Comparison of several stochastic and deterministic derivative-free global optimization algorithms. 	Lecture Notes in Computer Science  \textbf{11548}, 70--81 (2019)
 

### 2019-10-01 2019/Automation/Barkalov.tex
Fix GGergel

@@
 Р.Г.~СТРОНГИН, д-р~физ.-мат.~наук\\
-В.П.~ГГЕРГЕЛЬ, д-р~техн.~наук\\
+В.П.~ГЕРГЕЛЬ, д-р~техн.~наук\\
 К.А.~БАРКАЛОВ, канд.~физ.-мат.~наук\\

### 2019-10-07 2016/JoGO_GPU/BarkalovGergel.tex
JOGO2016 paper first version

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Journal of Global Optimization}
%
\begin{document}

\title{Parallel global optimization on GPU 
	\thanks{The research is supported by the grant of the Ministry of education and science of the Russian Federation (the agreement of August 27, 2013 No 02.B.49.21.0003)}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}
%\titlerunning{Short form of title}        % if too long for running head

\author{Konstantin Barkalov         \and
        Victor Gergel 
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{K. Barkalov  \at
              Computing Mathematics and Cybernetics Faculty, Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
              Tel.: +7-831-4623356\\
              Fax: +7-831-4623085\\
              \email{barkalov@vmk.unn.ru} 
           \and
           V. Gergel \at
							\email{gergel@unn.ru} 
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
This work considers a parallel algorithm for solving multidimensional multiextremal optimization problems. This algorithm uses Peano-type space filling curves for dimension reduction. Conditions of non-redundant parallelization of the algorithm are considered. Implementation of the algorithm on modern computing systems with the use of a graphics processing unit (GPU) is investigated. Speedup of the algorithm using GPU as compared with the same algorithm implemented on CPU only is demonstrated experimentally. Computing experiments are carried out on a series of several hundred multidimensional multiextremal problems.
\keywords{Global optimization \and Parallel computing \and Graphics processing unit \and Peano-type space filling curves \and Characteristical algorithms}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction} \label{intro}

This paper considers parallel algorithms for solving multiextremal optimization problems. The objective function in such problems is characterized by existence of several local optima (usually their number is not known and can be large). In multiextremal problems the possibility of a reliable estimation of the global optimum is essentially based on the availability of an a priori information about the function, allowing us to link possible values of an optimizable function with known values at search points. Very often, such information about the problem is presented as the assumption that the objective function $\varphi$ satisfies a Lipschitz condition with an a priori unknown constant $L$. The objective function can be represented by the algorithm for calculating its values in the search domain, i.e. it can be a ``black-box'' function. Similar problems are common in applications (problems of optimal design of objects and technological processes in various fields of technology, problems of model fitting according to observed data in scientific research, optimization of control processes, etc., see the review \cite{RefPinter}).

Multiextremal optimization problems offer little in terms of analytical solution, while numerical solution is computationally challenging as computational costs often grow exponentially in dimension. By employing modern parallel computational systems it is possible to broaden the area of applicability of global optimization methods. At the same time, the problem of efficient parallelization of the global search process arises. Therefore, development of efficient parallel methods for numerical solution of multiextremal optimization problems and their software implementation for modern computational systems is the obvious focus of research in this area.

One of the promising trends in the field of parallel global optimization (as in many areas connected with software implementation of complex algorithms) is the use of graphics processing units (GPUs). During the last decade, productivity of GPUs has been rapidly growing to satisfy the continuously growing demand of graphic application developers. Furthermore, some principles of development of graphics hardware have changed over the past few years. Today a GPU is a high-performance flexibly programmable and massively parallel processor that can provide solution to many computationally complex problems \cite{RefHwu}.

However, in the field of global optimization the potential of GPUs is not completely utilized. GPUs are generally used for parallelization of algorithms, which one way or another are based on the concept of random search (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi}); a review of this trend of GPU optimization algorithms is given in \cite{RefLangdon}. Owing to their stochastic nature, algorithms of this type generally do not guarantee convergence to a globally optimal solution, and that sets them apart from deterministic methods.

Some parallel variants are proposed for many deterministic algorithms of lipschitzian global optimization with guaranteed convergence \cite{RefGergel2005,RefEvtushenko,RefHe,RefPaulavicius}. However, these versions of algorithms are parallelized on CPU using MPI and/or OpenMP technologies; there are no implementations on GPU. The present article contains the results of an investigation into the GPU implementation of the global search parallel algorithm developed according to the information-statistical approach presented in the monograph \cite{RefStrongin2000}.

The main part of the paper has the following structure. Section \ref{sec:2} contains a description of the basic algorithm of global search. Section \ref{sec:3} contains a scheme of its parallelization, and section \ref{sec:4} includes theoretical statements that characterize the speedup and non-redundancy of the parallel algorithm. In section \ref{sec:5}, ways of possible implementation of the parallel algorithm on GPU are discussed. Section \ref{sec:6} contains the results of numerical experiments; a comparison of the sequential algorithm and parallel CPU and GPU versions for solving a series of test multidimensional multiextremal problems is also made. Section \ref{sec:7} concludes the paper.

\section{Multidimensional global search algorithm} \label{sec:2}

Let us consider the problem of search for a global minimum of an $N$-dimensional function $\varphi(y)$ within a hyper interval $D$
\begin{eqnarray}\label{eq:1}
& \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
& D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
\end{eqnarray}

Let us assume that the function $\varphi$ satisfies a Lipschitz condition with an a priori unknown constant $L$
\[
\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D,\; 0<L<\infty.
\]

There is a number of ways to adapt effective one-dimensional algorithms for solving multidimensional problems; see, for example, the diagonal partitions method in \cite{RefSergeyev2006} or the simplicial partitions method in \cite{RefZilinskas}. In this paper, we will use the approach based on the idea of dimension reduction by means of a Peano curve $y(x)$, which continuously and unambiguously maps the unit interval [0,1] onto the $n$-dimensional cube
\[
\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq N\right\}=\left\{y(x):0\leq x \leq 1 \right\}.
\]
Problems of numerical construction of Peano-type space filling curves and the corresponding theory are considered in detail in \cite{RefStrongin2000}, \cite{RefSergeyev2013}. Here we will note that a numerically constructed curve (\textit{evolvent}) is an approximation to a theoretical Peano curve with an accuracy $2^{-m}$ , where $m$ is an evolvent construction parameter. It means that using an evolvent we can distinguish points in a multidimensional space, the coordinates of which differ by not less than $2^{-m}$; closer points are considered as coincident. Thereby an evolvent constructed with a fixed $m$ allows obtaining a solution of a problem with an accuracy of not more than $2^{-m}$ by coordinate.

By using this kind of mapping it is possible to reduce the multidimensional problem~(\ref{eq:1}) to a one-dimensional problem
\[
\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}}.
\]
An important property of such mapping is preservation of boundedness of function relative differences  (see \cite{RefStrongin2000}): if the function $\varphi(y)$ in the domain $D$ satisfies the Lipschitz condition, then the function $\varphi(y(x))$ on the interval $[0,1]$ will satisfy a uniform H{\"o}lder condition
\[
\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N},
\]
where the H{\"o}lder constant $H$ is linked to the Lipschitz constant $L$ by the relation
\[
H=4Ld\sqrt{N},\;\; d=\max{\left\{b_i-a_i:1\leq i \leq N\right\}}.
\]

Therefore, it is possible, without limitation of generality, to consider minimization of one-dimensional function
\[
f(x)=\varphi(y(x)), \;\; x\in[0,1],
\]
satisfying the H{\"o}lder condition.

The considered algorithm for solving this problem (here, according to \cite{RefStrongin2000}) involves constructing a sequence of points $x^k$, where the values of minimized function $z^k = f(x^k)$ are calculated. Let us call the process of calculating the function value (including the construction of an image $y^k=y(x^k)$) the ``trial'', and the pair $(x^k, z^k)$, the ``trial result''. The set of pairs $\left\{(x^k, z^k)\right\}, 1\leq k\leq n,$ makes up the search data collected using the method after carrying out $n$ steps. The rules that determine the work of the global search algorithm are as follows.

At the first iteration of the method the trial is carried out at an arbitrary internal point $x^1$ of the interval $[0,1]$. The point of trial at the next iteration $(k+1)$ is determined according to these rules.

Step 1. Renumber points of the set
\[
X_k=\{x^1,\dots,x^k\}\cup\left\{0\right\}\cup\left\{1\right\},
\]
which includes boundary points of the interval $[0,1]$ and the points of the previous trials, with subscripts in increasing order of coordinate values, i.e.
\[
0=x_0<x_1<\dots <x_k<x_{k+1}=1.
\]

Step 2. Supposing that  $z_i=f(x_i), \; 1\leq i \leq k$, calculate values 
\begin{equation}\label{eq:11}
\mu = \max_{2\leq i \leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
\end{equation}
\[
M = \left\{
   \begin{array}{lr}
     r\mu, & \mu > 0,\\
     1, & \mu = 0,
   \end{array}
\right.
 \]
where $r>1$ is a preset parameter of the method, and $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$.

Step 3. Calculate a \textit{characteristic} for every interval $(x_{i-1}, x_i), \; 1\leq i \leq k+1,$   according to the following formulas
\[
R(1)=2\Delta_1-4\frac{z_1}{M},
\]
\begin{equation}\label{eq:14}
R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\frac{z_i+z_{i-1}}{M},1<i<k+1,
\end{equation}
\[
R(k+1)=2\Delta_{k+1}-4\frac{z_k}{M}.
\]

Step 4. Find interval $(x_{t-1},x_t)$ with the maximum characteristic
\[
R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
\]

Step 5. Carry out a trial at the point $x^{k+1}\in(x_{t-1},x_t)$, calculated using the following formulas
\[
x^{k+1} = \frac{x_t+x_{t-1}}{2}, \textrm{ if } t=1 \textrm{ or } t=k+1,
\]
\[
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N, \textrm{ if } 1<t<k+1.
\]

The algorithm terminates if the condition $\Delta_t<\epsilon$ is satisfied; here $\epsilon>0$ is the preset accuracy. For estimation of the global solution, values
\[
f_k^\ast=\min_{1\leq i \leq k}f(x^i), \ x_k^\ast=\arg \min_{1\leq i \leq k}f(x^i).
\]
are selected.

Remark 1. The normalized characteristics $R(i)$ from (\ref{eq:14}) can be viewed as probabilities of locating a global minimum within the interval $(x_{i-1},x_i), 1\leq i \leq k+1$. This statement is proved in \cite{RefStrongin2000}. Thus, at each iteration a new trial point is selected inside the interval, which has the greatest probability of finding a global minimum.

Remark 2. The proof of convergence of this algorithm is provided in \cite{RefStrongin2000}. The modifications taking into account existence of inequality constraints in the problem and the information about the objective function derivative are given in \cite{RefBarkalov,RefGergel1996,RefGergel1997}.

Let us illustrate the work of this \textit{global search algorithm} (GSA) using a dimension reduction for minimization of a two-dimensional multiextremal function
\begin{eqnarray} \nonumber \label{eq:19}
\varphi(y)= -&\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2+\right. \\
&\left.\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},\\ \nonumber
\end{eqnarray}
where
\begin{eqnarray} \nonumber
& y=(y_1,y_2)\in R^2, 0 \leq y_s \leq 1, s=1,2, \\ \nonumber
& g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2),  \\ \nonumber
& h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2), \nonumber 
\end{eqnarray}
and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$. The experiment used the method parameters $r=3$, $\epsilon=10^{-3}$, and the evolvent construction parameter $m=12$. Fig.~\ref{fig:1} shows the contour lines of a function generated according to (\ref{eq:19}) and the points of 1103 trials carried out using the method before the required accuracy was obtained.
\begin{figure}
	\center
  \includegraphics[width=0.75\textwidth]{fig1.jpg} 
  \caption{Minimization of a two-dimensional function by the sequential algorithm}
  \label{fig:1}       % Give a unique label
\end{figure}

\section{Synchronous parallel algorithm} \label{sec:3}

Let us now consider $p>1$ computing elements (cores in CPU or GPU). In this case, search process parallelization can be organized in several ways.

The first option is to parallelize calculation of the target function describing the optimized object. This method can provide some speedup, but is specific for each problem.

The second option is to parallelize implementation of algorithm computing rules for selection of the next trial point. In this case the way of parallelization will depend on the class of algorithms. Moreover, these rules are often rather simple and it is inexpedient to parallelize them (overhead costs of the organization of parallelism will reduce possible speedup to zero).

The third option is to  change the scheme of an algorithm for the purpose of carrying out several trials in parallel. This approach is most promising since it is characterized by efficiency (the part of computing process, in which the bulk of calculations is carried out, is parallelized) and generality (it is applicable for a wide class of characteristical algorithms of multiextremal optimization \cite{RefGrishagin1997}).

Let us describe a change of rules of global search algorithm for the purpose of simultaneous (synchronous) carrying out in one iteration of method $p$ of trials. Let us make $k(n)$ the total number of trials, carried out after $n$ parallel iterations.

Suppose $n\geq 1$  iterations of the method are carried out, in which trials were performed at $k=k(n)$ points $x^i,1\leq i \leq k$. Then points $x^{k+1},\dots,x^{k+p}$  of search trials of the next $n+1$ iteration are determined according to the rules.

Steps 1--3 of the parallel algorithm fully repeat steps 1--3 of the sequential global search algorithm.

Step 4. Arrange characteristics  $R(i), 1 \leq i \leq k+1$, in decreasing order 
\begin{equation}\label{eq:21}
R(t_1)\geq R(t_2)\geq \dots \geq R(t_{k}) \geq R(t_{k+1})
\end{equation}
and select $p$ maximum characteristics with interval numbers $t_j, 1\leq j \leq p$.

Step 5. Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), 1\leq j\leq p$, calculated using the formulas
\[
x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2}, \textrm{ if } t_j=1 \textrm{ or } t_j=k+1,
\]
\[
x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - \mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N, \textrm{ if } 1<t_j<k+1.
\]

The algorithm terminates if the condition $\Delta_{t_j}<\epsilon$ is satisfied at least for one number $t_j, 1 \leq j \leq p$ ; $\epsilon>0$ is the preset accuracy.  For estimation of the global solution, values
\[
f_k^\ast=\min_{1\leq i \leq k}f(x^i), \ x_k^\ast=\arg \min_{1\leq i \leq k}f(x^i).
\]
are selected.

The modifications of the parallel algorithm with the account of the information about the minimized function derivative are given in \cite{RefGergel1999}.

This method of organizing parallel computing has the following justification \cite{RefStrongin2000,RefGrishagin1997}. The characteristics of intervals (\ref{eq:14}) used in the global search algorithm can be considered as probability measures of the global minimum point location in these intervals. Inequalities (\ref{eq:21}) arrange intervals according to their characteristics, and trials are carried out in parallel in the first $p$ intervals with the largest probabilities.

Let us use a parallel algorithm of global search for solving a problem from section \ref{sec:2} with the numbers of parallel trials $p=2$ and $p=4$. With $p=2$ the number of iterations was $569$, and the number of trials was $1138$. With $p=4$ the number of iterations was reduced to $270$, and $1080$ trials were carried out. Fig.~\ref{fig:2} shows contour lines of the function with points of the trials performed by the method with $p=2$. Points of redundant trials carried out using a parallel algorithm are marked in the figure by "+". The concept of redundancy is discussed in more detail in the following section.
\begin{figure}
	\center
  \includegraphics[width=0.75\textwidth]{fig2.jpg} 
  \caption{Minimization of a two-dimensional function by the parallel algorithm}
  \label{fig:2}
\end{figure}

\section{Speedup and non-redundancy of a parallel algorithm} \label{sec:4}

Let us describe theoretical properties of a parallel algorithm, which characterize its speedup. One of the main indicators of efficiency of parallel algorithms (in any field, not only in global optimization) is a speedup in time
\[
S(p)=T(1)/T(p),
\]
where $T(1)$ is the time required to solve a problem by a sequential algorithm, and $T(p)$ is the time for solving the same problem by a parallel algorithm in a system with $p$ computing elements. The characteristic of efficiency of a parallel algorithms (in relation to optimization algorithms) is also a speedup in iterations
\begin{equation}\label{eq:26}
s(p)=n(1)p/n(p),
\end{equation}
where $n(1)$ is the number of trials carried out using the sequential method, and $n(p)$ is the number of trials carried out using the parallel method with $p$ processors. This characteristic is especially important since for solving applied problems the time required to carry out a trial exceeds the time required to process its results.

It is obvious that the number of trials $n(p)$ for sequential and parallel algorithms described in sections \ref{sec:2} and \ref{sec:3} will differ. Actually, the sequential algorithm of global search when selecting point $x^{k+1}$ of the next $k+1$ trial possesses complete information received at the previous $k$ iterations. Based on the same information, the parallel algorithm of global search selects $p$ points $x^{k+j}, 1\leq j \leq p$, rather than one at iteration $k+1$. It means that selection of point $x^{k+j}$ is carried out in absence of information on the trial results at points $x^{k+i}, 1\leq i<j$. Only the first point $x^{k+1}$ will coincide with the point selected by the sequential algorithm. Other trials points generally cannot coincide with the points generated by the sequential algorithm. The efficiency offered by the use of parallel processors can be reduced when such trials are performed. Therefore, let us consider such trials as ``redundant'', and the value 
\[
\lambda(p) = \left\{
   \begin{array}{lr}
     (n(p)-n(1))/n(p), & n(p) > n(1),\\
     0, & n(p)\leq n(1),
   \end{array}
\right.
\]
will represent ``method redundancy''.

Let us set a series of trials $\left\{x^k\right\}$ and $\left\{y^m\right\}$ generated correspondingly by sequential and parallel algorithms for solving the same problem with $\epsilon = 0$ in the stopping condition. The following theorem from \cite{RefStrongin2000} determines the number of computing elements $p$, which can be involved for non-redundant parallelization.

\textbf{Theorem}. Suppose $x^\ast$ is the point of global minimum, $x'$ is the point of the local minimum of function $f(x)$, and the following conditions are fulfilled:
\begin{enumerate}
	\item
	Inequality 
	\begin{equation}\label{eq:28}
	f(x')-f(x^\ast)\leq \delta, \delta>0,
	\end{equation}
	holds.
	\item
	The initial $q(l)$ trials of the sequential and parallel methods coincide, i.e.
	\[
	\left\{x^1,\dots,x^{q(l)}\right\}=\left\{y^1,\dots,y^{q(l)}\right\},
	\]
	where 
	\[
	\left\{x^1,\dots,x^{q(l)}\right\}\subset\left\{x^k\right\}, \ \left\{y^1,\dots,y^{q(l)}\right\}\subset\left\{y^m\right\}.
	\]
	\item
	There exists a point $y^n\in \left\{y^m\right\}, n<q(l),$ such that $x'\leq y^n \leq x^\ast$ or \mbox{$x^\ast \leq y^n \leq x'$}.
	\item
	For the value $M$ from (\ref{eq:11}) the following inequality
	\[
	M>2^{2-1/N}H
	\]
	holds, where $H$ is the H{\"o}lder constant of the minimized function.
\end{enumerate}

Then a parallel algorithm of global search using two processors will be non-redundant (i.e. $s(2)=2$, $\lambda(2)=0$) while the following condition is satisfied 
\begin{equation}\label{eq:32}
\left(x_{t_j}-x_{t_j-1}\right)^{1/N} > \frac{4\delta}{M-2^{2-1/N}H}, 1\leq j\leq p,	
\end{equation}
where $t_j$ are determined according to (\ref{eq:21}).

\textbf{Corollary}. Let the objective function $f(x)$ have $Q$ local minimum points $\left\{x'_1,\dots,x'_Q\right\}$, for which condition (\ref{eq:28}) is fulfilled, and let there exist trial points $y^{n_i}, 1\leq i \leq Q,$ such as
\[
y^{n_i}\in \left\{y^1,\dots,y^{q(l)}\right\},
\]
\[
\alpha_i \leq y^{n_i} \leq \alpha_{i+1}, \; \alpha_i, \alpha_{i+1} \in \left\{x^\ast, x'_1,\dots,x'_Q\right\}, \; 1\leq i \leq Q.
\]

Then, if the theorem conditions are satisfied, the parallel algorithm of global search with $Q+1$ processors will be non-redundant (i.e. $s(Q+1)= Q+1$, \mbox{$\lambda(Q+1)=0$}), while condition (\ref{eq:32}) is satisfied.

The theorem corollary plays a special role in solving multidimensional problems reduced to one-dimensional problems by means of the Peano-like evolvent $y(x)$. The evolvent $y(x)$, which is an approximation to the Peano curve, has the effect of ``splitting'' of a global minimum point $y^\ast\in D$ to several preimages in the interval $[0,1]$. If function $\varphi(y)$ has the only global minimum in $D$, the ``reduced'' function $f(x)$ can have up to $2^N$ local extremum points close (by value) to a global extremum point (see \cite{RefStrongin2000}). By applying a parallel global search algorithm for minimization of a similar function it is possible to achive non-redundancy when using up to $2^N+1$ computing elements.

The parallel algorithm of global search has demonstrated a good speedup in implementation on CPU in systems with a small number of processors. In \cite{RefGrishagin1997,RefStrongin2003} a speedup of algorithm close to linear was experimentally verified when carrying out 2-4 parallel trials. The results obtained with the parallel GSA for minimization of 100 two-dimensional multiextremal functions described by formula (\ref{eq:19}) are shown in table \ref{tab:1}. All experiments used method parameters $r=2.9$, and $\epsilon=10^{-3}$, and the global minimum was found with the required accuracy in all $100$ problems. The table shows the number of used processors $p$, the average number of trials $n(p)$ rounded to integers and average speedup in iterations $s(p)$ from (\ref{eq:26}).

\begin{table}
	\caption{Results of minimization of $100$ two-dimensional functions}
	\label{tab:1}
	\center
	\begin{tabular}{lll}
		\hline\noalign{\smallskip}
		 $p$ & $n(p)$ & $s(p)$ \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
			1 &	1575 &	\dots \\
			2 &	1596 &	1.97 \\
			3 &	1562 &	3.02 \\
			4 &	1599 &	3.94 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\section{Implementation on GPU} \label{sec:5}

What are the advantages of GPU over CPU when implementing global optimization algorithms? First, CPU has only a small (up to $16$ in the advanced models) number of cores working independently at a high clock frequency. Each CPU core is a powerful computer. GPU, on the contrary, works at low clock frequencies and uses hundreds of simpler computing elements. Second, a considerable part of a CPU crystal is occupied by a high-speed cache memory, while practically the whole of GPU consists of arithmetic and logic units. Therefore, GPUs are especially effective in problems, in which the number of arithmetic operations is great in comparison with memory operations.

With respect to methods of global optimization, parallel computing of many values of the objective function simultaneously is an operation which can be efficiently implemented on GPU. In the function value calculation procedure implemented on GPU transfers of information from CPU to GPU will be minimal: it is necessary only to transfer the coordinates of trial points to GPU and to receive the values of the function in these points. The functions that determine the processing of trial results according to algorithm and require working with a large volume of collected search information can be efficiently implemented on CPU.

A general scheme showing the organization of calculations using GPU is given in Fig.~\ref{fig:3}. According to this scheme, steps 1 -- 4 of a parallel global search algorithm are carried out on a CPU. The coordinates of $p$ points of trials calculated at step 4 of the algorithm are collected in an intermediate buffer, and then are transferred to the graphics processing unit. Calculation of function values at these points is carried out on a GPU, and then the results of trials (again via the intermediate buffer) are transferred to the CPU.

\begin{figure}
	\center
  \includegraphics[width=0.75\textwidth]{fig3.jpg} 
  \caption{Scheme of information exchanges in a GPU algorithm}
  \label{fig:3}
\end{figure}

The use of several hundred of parallel computing elements (as in modern GPU) might not give a speedup of the search process by hundreds of times. In this case, the conditions of the theorem of non-redundant parallelization can be violated: the number of local extremums will be less than the number of computing cores. Then (according to the theorem and its corollary) a parallel GSA will generate redundant trial points. Nevertheless, despite certain redundancy, the overall operating time of a GPU algorithm will be less than the operating time of a CPU algorithm. This has been verified by computing experiments, the results of which are given in section \ref{sec:6}.

\section{Results of numerical experiments} \label{sec:6}

Computing experiments were carried out on one of the nodes of a high-performance cluster of the State University of Nizhny Novgorod. The cluster node includes 2 Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM, and NVIDIA Tesla M2090 GPU. The CPU has 8 cores (i.e. 16 physical cores and 16 virtual cores in hyper-threading mode are available in the node). The graphics processing unit includes 16 multiprocessors (512 cores). For implementation of the GPU algorithm CUDA Toolkit 6.0 was used.

It should be noted that widely known test problems from the field of multidimensional global optimization are characterized by small time required to calculate the objective function value. Usually, such a calculation is nothing else than a summation of several (according to problem dimension) values of elementary functions. For example, the well known Rastrigin test function, the global minimum of which is located in $y^\ast = (0,\dots,0)$ with $\varphi(y^\ast)=0$, is set by the expression
\begin{equation}\label{eq:33}
\varphi(y)=10N+\sum^{N}_{i=1}\left[y^2_i-10\cos(2\pi y_i)\right], \; -2.2\leq y_i \leq 1.8, \; 1\leq i \leq N.
\end{equation}

Therefore, for the purpose of imitation of the computational complexity inherent to applied problems of optimization \cite{RefBarkalov2013}, calculation of the objective function in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (series summation from 20 thousand elements).

It is noteworthy that an important parameter for starting any GPU program (unlike a parallel program on CPU) is not only the number of threads $p$, but also the size of the \textit{thread block}. Separate threads on GPU are grouped in blocks of threads of identical size, thus each thread block is executed on a separate multiprocessor. Threads in a thread block can interact effectively using information in shared memory and synchronization. For the purpose of assessing the importance of the size of a thread block during the execution period of a GPU algorithm let us perform a preliminary series of experiments with the test function (\ref{eq:33}) with $N=5$. Let us vary the number $p$ of the used GPU threads and the size $pB$ of a thread block. All the other parameters of the method do not change ($r=2.3$, $\epsilon = 10^{-3}$, the evolvent construction parameter $m=10$). Table \ref{tab:2} shows the GPU algorithm operating time depending on $p$ and $pB$.
\begin{table}
	\caption{GPU algorithm operating time depending on thread block size}
	\label{tab:2}
	\center
	\begin{tabular}{llllll}
		\hline\noalign{\smallskip}
%		 $p$ & $pB=16$ & $pB=32$ & $pB=64$ & $pB=128$ & $pB=256$  \\
		$p$ & \multicolumn{5}{l}{ \hfil $pB$ \hfil } \\
		\noalign{\smallskip} \cline{2-6} \noalign{\smallskip}
		 & 16 & 32 & 64 & 128 & 256  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		100 &	39.43 &	41.5 &	39.71 &	39.85 &	39.61 \\
		300 &	24.84 &	24.52 &	24.54 &	24.66 &	25.08 \\
		500 &	9.25 &	9.45 &	9.24 &	9.4 &	9.3\\
		700 &	11.8 &	13.43 &	11.78 &	11.91 &	12.03\\
		900 &	10.9 &	10.91 &	10.85 &	10.86 &	10.96\\
		1000 &	10.92 &	10.8 &	10.76 &	10.9 &	10.74\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

As is seen from this table, there is a speedup of the algorithm in the case of an increase in the number of trials at one iteration up to the number of cores on GPU, with subsequent saturation. Owing to a weak information dependence the size of the thread block $pB$ does not have decisive effect, therefore let us use $pB=64$ in all subsequent experiments. When carrying out this set of experiments the load on GPU was 15 -- 85 \% (depending on the number of threads); the load was measured by GPUZ program.

After the parameters of the GPU algorithm have been determined let us make the main computing experiments on a series of multidimensional multiextremal problems. In work \cite{RefGaviano} a GKLS generator allowing generation of multiextremal optimization problems with known properties (number of local minima, size of their domains of attraction, global minimum point, etc.) is described.

The results of numerical comparison of the three sequential algorithms -- DIRECT \cite{RefJones}, DIRECT\textit{l} \cite{RefGablonsky} and global search algorithm (GSA) from section \ref{sec:2} -- are provided below (results of the work of the first two algorithms are given in paper \cite{RefGaviano}). Numerical comparison was carried out on function classes \textit{Simple} and \textit{Hard} of dimension 4 and 5 from \cite{RefGaviano} since solving a problems of dimension 2 and 3 demands a small number of iterations and it is impractical to use GPU for solving these problems. Global minimum $y^\ast$ was considered to be found, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e. $\left\|y^k-y^\ast\right\|\leq\delta$. The size of the vicinity was selected (according to \cite{RefGaviano}) as $\delta = \left\|b-a\right\|\sqrt[N]{\Delta}$, $N$ -- problem dimension, $a$ and $b$ -- borders of search domain $D$, parameter $\Delta=10^{-6}$ at $N=4$ and $\Delta=10^{-7}$ at $N=5$. When using the GSA for the class \textit{Simple} parameter $r=4.5$ was selected, for the class \textit{Hard} $r=5.6$ was selected; evolvent construction parameter was fixed as $m=10$. The maximum allowable number of iterations was $K_{max} = 10^6$.

The average number of iterations $k_{av}$ performed by the method for solving a series of problems from these classes is shown in table \ref{tab:3}. Symbol ''$>$''  reflects a situation in which not all problems of the class were solved by the method. This means that the algorithm was stopped as the maximum allowable number of iterations $K_{max}$ was achieved. In this case, value $K_{max}=10^6$ was used for calculating the average value of the number of iterations $k_{av}$ that corresponds to the lower estimate of this average value. The number of unsolved problems is specified in brackets.

\begin{table}
	\caption{Average number of iterations $k_{av}$}
	\label{tab:3}
	\center
	\begin{tabular}{lllll}
		\hline\noalign{\smallskip}
  	 $N$ & Problem class & DIRECT & DIRECT\textit{l} & GSA \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
			4 &	\textit{Simple}	& $>$47282(4) &	18983 &	11953 \\
	      & \textit{Hard} &	$>$95708(7) &	68754 &	25263 \\
			5	& \textit{Simple} &	$>$16057(1) &	16758 &	15920 \\
				& \textit{Hard} &	$>$217215(16) &	$>$269064(4) & $>$148342(4) \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

Table \ref{tab:3} shows that GSA outperforms DIRECT and DIRECT\textit{l} methods on all classes of problems in terms of average number of iterations. In class 5-\textit{Hard} all the three methods failed to solve all problems: DIRECT did not solve 16 problems, DIRECT\textit{l} and GSA, 4 problems.

Let us estimate now the speedup when using the parallel global search algorithm implemented on CPU depending on the number $p$ of the used cores. Table \ref{tab:4} shows the average number of iterations performed by the method when solving a series of problems, and table \ref{tab:5} shows the average time required to solve one problem. Tables \ref{tab:6} and \ref{tab:7} show speedup in time $S(p)$ and iterations $s(p)$, respectively; speedup of the parallel algorithm was measured in relation to the sequential one ($p=1$).

\begin{table}
	\caption{Average number of iterations on CPU}
	\label{tab:4}
	\center
	\begin{tabular}{llllll}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{l}{ $N=4$ } & & \multicolumn{2}{l}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		1 &	11953 &	25263 & &	15920 &	$>$148342(4) \\
		2 &	6014 &	16509 & & 17797 &	98667 \\
		4 &	2547 &	8970 & &	8246 &	45410 \\
		8 &	1413 &	4666 & &	3773 &	26732 \\
		16 &	756 &	1910 & &	1713 &	11322 \\
		32 &	323 &	1252 & &	781 &	11745 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Average problem solution time on CPU}
	\label{tab:5}
	\center
	\begin{tabular}{llllll}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{l}{ $N=4$ } & & \multicolumn{2}{l}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		1 &	3.02 &	6.05 & &	4.84 &	$>$42.5 \\
		2 &	1.8 &	4.48 & &	5.84 &	33.77 \\
		4 &	1.0 &	2.94 & &	3.18 &	15.62 \\
		8 &	0.71 &	1.88 & &	1.82 &	14.16 \\
		16 &	0.57 &	1.36 & &	1.44 &	8.37 \\
		32 &	0.38 &	1.25 & &	0.95 &	11.67 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Time speedup $S(p)$ on CPU}
	\label{tab:6}
	\center
	\begin{tabular}{llllll}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{l}{ $N=4$ } & & \multicolumn{2}{l}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		2 &	1.68 &	1.35 & &	0.83 &	$>$1.26 \\
		4 &	3.04 &	2.06 & &	1.52 &	2.72 \\
		8 &	4.27 &	3.22 & &	2.66 &	3.64 \\
		16 &	5.29 &	4.46 & &	3.36 &	5.08 \\
		32 &	8.04 &	4.85 & &	5.11 &	3.0 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Iteration speedup $s(p)$ on CPU}
	\label{tab:7}
	\center
	\begin{tabular}{llllll}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{l}{ $N=4$ } & & \multicolumn{2}{l}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		2 &	1.99 & 1.53 & &	0.89 &	$>$1.50 \\
		4 &	4.69 &	2.82 & &	1.93 &	3.27 \\
		8 &	8.46 &	5.41 & &	4.22 &	5.55 \\
		16 &	15.82 &	13.23 & &	9.29 &	13.1 \\
		32 &	37.04 &	20.18 & &	20.39 &	12.63 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Redundancy $\lambda(p)$ on CPU}
	\label{tab:8}
	\center
	\begin{tabular}{llllll}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{l}{ $N=4$ } & & \multicolumn{2}{l}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
2 &	0.01 & 0.23 & &	0.55 &	$>$0.25 \\
4 &	0 &	0.3 & &	0.52	& 0.18 \\
8 &	0 &	0.32 & &	0.47  &	0.31 \\
16 &	0.01 &	0.17& &	0.42  &	0.18 \\
32 &	0 &	0.37 & &	0.36  &	0.61 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

The data show five-fold time speedup, when using 32 cores for all classes (except 5-\textit{Hard}). The lack of expected speedup in class 5-\textit{Hard} in transition from 16 to 32 cores is explained by an increase in the average number of iterations and by the fact that out of 32 cores 16 are physical, and other 16 are virtual in hyper-threading mode, i.e. a real gain in performance in transition from 16 to 32 cores is less than twofold.

Let us make a similar experiment using a GPU version of the algorithm. Also let us fix the average number of iterations and average time required to solve one problem (see tables \ref{tab:last4} -- \ref{tab:last}). Let us measure the speedup and redundancy of GPU algorithm in relation to the CPU algorithm started with 32 cores (see column $p=32$ in tables \ref{tab:4} and \ref{tab:5}). The number of the used threads $p$ on GPU varied, all other parameters of the method did not change. The number of threads, which is greater, than the number of cores, was chosen to show the following effect: the number of iterations executed by the algorithm will be reduced in case of an increase of the threads number, while the algorithm operation time will not be reduced. The minimum time of operation is reached, if the number of used threads does not surpass the number of GPU cores.

\begin{table}
	\caption{Average number of iterations on GPU}
	\label{tab:last4}
	\center
	\begin{tabular}{llllll}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{l}{ $N=4$ } & & \multicolumn{2}{l}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & Simple & Hard & & Simple & Hard  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}		
100 &	105 &	352 & &	193  &	2452 \\
200 &	56 &	181 & &	105  &	809 \\
300 &	53 &	107 & & 62  &	508 \\
400 &	28 &	87 & &	44  &	376 \\
500 &	24 &	84 & &	41  &	353 \\
800 &	18 &	37 & &	55 &	246 \\
1000 &	12 &	33 & &	36 &	172 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Average problem solution time on GPU}
	\label{tab:last3}
	\center
	\begin{tabular}{llllll}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{l}{ $N=4$ } & & \multicolumn{2}{l}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & Simple & Hard & & Simple & Hard  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}		
		100  & 0.16 &	0.5 & &	0.33 & 3.46 \\
		200 &	0.16 &	0.47 & &	0.32  &	2.11 \\
		300 &	0.22 &	0.43 & &	0.29  &	1.98 \\
		400 &	0.16 &	0.44 & &	0.27  &	1.97 \\
		500 &	0.16 &	0.52 & &	0.31 &	2.25 \\
		800 &	0.19 &	0.37 & &	0.6 &	2.54 \\
		1000 &	0.15 &	0.41 & &	0.48  &	2.15 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Time speedup $S(p)$ on GPU}
	\label{tab:last2}
	\center
	\begin{tabular}{llllll}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{l}{ $N=4$ } & & \multicolumn{2}{l}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & Simple & Hard & & Simple & Hard  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}		
		100 &	2.3  &	2.49 & &	2.9  &	4.09 \\
		200 &	2.35 &	2.68 & &	2.99 &	6.72 \\
		300 &	1.75 &	2.93 & &	3.32 &	7.17 \\
		400 &	2.37 &	2.85 & &	3.58 &	7.19 \\
		500 &	2.38 &	2.41 & &	3.05 &	6.3 \\
		800 &	1.94 &	3.38 & &	1.59 &	5.56 \\
		1000 & 2.45 &	3.07 & &	1.98 &	6.59 \\				
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Iteration speedup $s(p)$ on GPU}
	\label{tab:last1}
	\center
	\begin{tabular}{llllll}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{l}{ $N=4$ } & & \multicolumn{2}{l}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & Simple & Hard & & Simple & Hard  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		100	&	3.07	&	3.55	& &	4.05	&	4.79	\\
		200	&	5.78	&	6.90	& &	7.45	&	14.51	\\
		300	&	6.11	&	11.70	& &	12.57	&	23.12	\\
		400	&	11.43	&	14.38	& &	17.80	&	31.21	\\
		500	&	13.54	&	14.98	& &	18.83	&	33.26	\\
		800	&	17.96	&	33.46	& &	14.19	&	47.68	\\
		1000	&	27.42	&	37.40	& &	21.91	&	68.21	\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Redundancy $\lambda(p)$ on GPU}
	\label{tab:last}
	\center
	\begin{tabular}{llllll}
		\hline\noalign{\smallskip}
		$p$ & \multicolumn{2}{l}{ $N=4$ } & & \multicolumn{2}{l}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & Simple & Hard & & Simple & Hard  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		100 & 0.02 & 0.0 & & 0.0 & 0.0  \\
		200 & 0.08 & 0.0 & & 0.0 & 0.0  \\
		300 & 0.35 & 0.0 & & 0.0 & 0.0  \\
		400 & 0.09 & 0.0 & & 0.0 & 0.0  \\
		500 & 0.13 & 0.04 & & 0.0 & 0.0  \\
		800 & 0.28 & 0.0 & & 0.43 & 0.0  \\
		1000 & 0.12 & 0.0 & & 0.3 & 0.0  \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

The results of our experiments show some obvious advantages of the GPU algorithm in comparison with the CPU algorithm. The algorithm using the maximum quantity of cores of CPU works slower than any of the given GPU implementations. In addition, the results show a considerable speedup in the number of iterations. For example, on average only 353 parallel iterations on GPU were sufficient to solve problems of the most difficult class considered -- 5-\textit{Hard}, whereas even with the use of all computing CPU cores over eleven thousand iterations were required. Finally, the GPU algorithm is almost non-redundant in relation to the CPU algorithm.

\section{Conclusions} \label{sec:7}

By now, parallel implementations have been suggested for the most of the existing sequential algorithms. A review of modern methods of parallel global optimization can be found, for example, in \cite{RefApuzzo}. The methods described in \cite{RefApuzzo} can be divided into two classes: deterministic methods of global optimization, and heuristic methods, which are one way or another based on the random search concept. Algorithms belonging to the latter class allow use of massive parallelism and are easily implemented on GPU (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi} and the review \cite{RefLangdon}).

Implementations on GPU for deterministic algorithms of global optimization has not been suggested yet. For example, parallelization of algorithms based on the ideas of the branch and bound method is carried out for a small number of processors using OpenMP and MPI \cite{RefPaulavicius}. It is noteworthy that the branch and bound method is applied in other areas of optimization; in particular, combinatorial optimization and integer programming. Using GPU shows good results in these areas (see \cite{RefBoukedjar}, \cite{RefCarneiro}). It is explained by the following factor. Problems of this class (such as the Travelling Salesman Problem or the Knapsack problem) are characterized by short time of criterion value calculation in one point and, at the same time, relatively long time required to process the results. Therefore, the most labor-intensive operations here are branching and bound computation, which are implemented on GPU (see \cite{RefBoukedjar}, \cite{RefCarneiro}).

This paper considers the problems of lipschitzian global optimization that are frequently encountered in applications. They are characterized by a considerable time required to calculate the function value in one point (in comparison with the processing of the results of such calculation). Therefore, calculation of the value of an optimized function in one point is done using a GPU, and the main objective of the algorithm (working at CPU) is to efficiently select points of parallel trials. This is carried out in the parallel implementation of GSA on GPU.

Summing up the results, let us note that use of graphics processing units for solving global optimization problems is a new, promising line of research as high efficiency of modern supercomputers is reached (generally) due to the use of graphics accelerators.

This paper considers a parallel algorithm for solving multidimensional multiextremal optimization problems and its implementation on GPU. This algorithm was developed within the framework of information-statistical approach. Theoretical information on the properties of the parallel algorithm characterizing the conditions of non-redundant parallelization are provided.

For the purpose of experimental demonstration of theoretical properties of the considered parallel algorithms, computing experiments on a series of several hundred test problems of different dimensions were performed. The results obtained with the sequential global search algorithm show its superiority over other known methods of global optimization. At the same time the parallel algorithm shows a good speedup and low redundancy (both GPU and CPU version), and the GPU version of the parallel algorithm shows a speedup of 2 to 6 times in comparison with the fastest parallel CPU version.

\begin{acknowledgements}

The authors of the paper express their gratitude to Ilya Lebedev for assistance in carrying out the computing experiments.

\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

% Non-BibTeX users please use
\begin{thebibliography}{}

\bibitem{RefPinter}
Pinter, J.D. (ed.): Global Optimization: Scientific and Engineering Case Studies. Springer (2006).

\bibitem{RefHwu}
Hwu, W.: GPU Computing Gems Emerald Edition (Applications of GPU Computing Series). Morgan Kaufmann (2011).

\bibitem{RefFerreiro}
Ferreiro, A.M., Garcia, J.A., Lopez-Salas, J.G., Vazquez, C.: An efficient implementation of parallel simulated annealing algorithm in GPUs. J. Glob. Optim. 57(3), 863--890 (2013)

\bibitem{RefZhu}
Zhu, W.: Massively parallel differential evolution--pattern search optimization with graphics hardware acceleration: an investigation on bound constrained optimization problems. J. Glob. Optim. 50(3), 417--437 (2011).

\bibitem{RefGarcia}
Garcia-Martinez, J.M., Garzon, E.M., Ortigosa, P.M.: A GPU implementation of a hybrid evolutionary algorithm: GPuEGO. J. Supercomput. (2014) doi:10.1007/s11227-014-1136-7

\bibitem{RefMussi}
Mussi, L., et al.: GPU implementation of a road sign detector based on particle swarm optimization. Evol. Intel. 3 (3), 155--169 (2010)

\bibitem{RefLangdon}
Langdon, W.B.: Graphics processing units and genetic programming: an overview. Soft. Comput. 15(8), 1657--1669 (2011)

\bibitem{RefGergel2005}
Gergel, V.P., Strongin, R.G.: Parallel computing for globally optimal decision making on cluster systems. Future Generation Computer Systems, 21(5), 673--678 (2005)

\bibitem{RefEvtushenko}
Evtushenko, Yu.G., Malkova, V.U., Stanevichyus, A.A.: Parallel global optimization of functions of several variables. Computational Mathematics and Mathematical Physics, 49(2), 246--260 (2009)

\bibitem{RefHe}
He, J., Verstak, A., Watson, L.T., Sosonkina, M.: Design and implementation of a massively parallel version of DIRECT. Comput. Optim. Appl. 40(2), 217--245 (2008)

\bibitem{RefPaulavicius}
Paulavicius, R., Zilinskas, J., Grothey, A.: Parallel branch and bound for global optimization with combination of Lipschitz bounds. Optimization Methods \& Software 26(3), 487--498, (2011)

\bibitem{RefStrongin2000}
Strongin, R.G., Sergeyev, Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000)

\bibitem{RefBarkalov}
Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Computational Mathematics and Mathematical Physics 42(9), 1289--1300 (2002)

\bibitem{RefGergel1996}
Gergel, V.P.: A method of using derivatives in the minimization of multiextremum functions. Computational Mathematics and Mathematical Physics 36(6), 729--742 (1996)

\bibitem{RefGergel1997}
Gergel, V.P.: A global optimization algorithm for multivariate functions with lipschitzian first derivatives. J. Glob. Optim.10(3), 257--281 (1997)

\bibitem{RefGergel1999}
Gergel, V.P., Sergeyev, Ya.D.: Sequential and parallel algorithms for global minimizing functions with lipschitzian derivatives. Computers and Mathematics with Applications, 37(4-5), 163--179 (1999)

\bibitem{RefSergeyev1994}
Sergeyev, Ya.D., Grishagin, V.A.: A parallel method for finding the global minimum of univariate functions. J. Optim. Theory Appl. 80(3), 513--536 (1994)

\bibitem{RefSergeyev2006}
Sergeyev Ya.D., Kvasov D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim. 16(3), 910--937 (2006)

\bibitem{RefZilinskas}
Zilinskas, J.: Branch and bound with simplicial partitions for global optimization. Mathematical Modelling and Analysis 13(1), 145--159 (2008)

\bibitem{RefGrishagin1997}
Grishagin, V.A., Sergeyev, Ya.D., Strongin, R.G.: Parallel characteristical algorithms for solving problems of global optimization. J. Glob. Optim. 10(2), 185--206 (1997)

\bibitem{RefStrongin2003}
Strongin, R.G., Sergeyev, Ya.D.: Global optimization: fractal approach and non-redundant parallelism. J. Glob. Optim. 27(1), 25--50 (2003)

\bibitem{RefSergeyev2013}
Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013)

\bibitem{RefBarkalov2013}
Barkalov, K., Polovinkin, A., Meyerov, I., Sidorov, S., Zolotykh, N.: SVM regression parameters optimization using parallel global search algorithm // Lecture Notes in Computer Science. 7979, 154--166 (2013)

\bibitem{RefGaviano}
Gaviano, M., Lera, D., Kvasov, D.E., Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Trans. Math. Software 29, 469--480 (2003)

\bibitem{RefJones}
Jones, D.R., Perttunen, C.D., Stuckman, B.E.: Lipschitzian optimization without the lipschitz constant. J. Optim. Theory Appl. 79(1), 157--181 (1993)

\bibitem{RefGablonsky}
Gablonsky, J.M., Kelley, C.T.: A locally-biased form of the DIRECT algorithm. J. Glob. Optim., 21(1), 27--37 (2001)

\bibitem{RefApuzzo}
D'Apuzzo, M.,  Marino, M., Migdalas, A., Pardalos, P.M., Toraldo, G.: Parallel computing in global optimization. In Handbook of Parallel Computing and Statistics, Chapman \& Hall, London, 225--258 (2006)

\bibitem{RefBoukedjar}
Boukedjar A., Lalami M. E., El Baz D.: Parallel Branch and Bound on a CPU-GPU System. In 20th Int. Conf. on Parallel,  Distributed and Network-Based Processing, 392--398 (2012)

\bibitem{RefCarneiro}
Carneiro, T. et al.: A new parallel schema for branch-and-bound algorithms using GPGPU. In 23rd Int. Symposium on Computer Architecture and High Performance Computing, 41--47 (2011)

\end{thebibliography}

\end{document}
% end of file template.tex

### 2019-10-07 2016/JoGO_GPU/BarkalovGergel.tex
JOGO2016 paper second version

@@
 \begin{abstract}
-This work considers a parallel algorithm for solving multidimensional multiextremal optimization problems.
+This work considers a parallel algorithm of solution of multidimensional multiextremal optimization problems.
 This algorithm uses Peano-type space filling curves for dimension reduction.
-Conditions of non-redundant parallelization of the algorithm are considered.
-Implementation of the algorithm on modern computing systems with the use of a graphics processing unit (GPU) is investigated.
+Conditions of non-redundant parallelization of the algorithm is considered.
+The issue of implementation of the algorithm on modern computing systems with use of a graphics processing unit (GPU) is investigated.
 Speedup of the algorithm using GPU as compared with the same algorithm implemented on CPU only is demonstrated experimentally.
-Computing experiments are carried out on a series of several hundred multidimensional multiextremal problems.
+Computing experiments are carried out on a series of several hundred of multidimensional multiextremal problems.
 \keywords{Global optimization \and Parallel computing \and Graphics processing unit \and Peano-type space filling curves \and Characteristical algorithms}
@@
 
-This paper considers parallel algorithms for solving multiextremal optimization problems.
-The objective function in such problems is characterized by existence of several local optima (usually their number is not known and can be large).
-In multiextremal problems the possibility of a reliable estimation of the global optimum is essentially based on the availability of an a priori information about the function, allowing us to link possible values of an optimizable function with known values at search points.
-Very often, such information about the problem is presented as the assumption that the objective function $\varphi$ satisfies a Lipschitz condition with an a priori unknown constant $L$.
-The objective function can be represented by the algorithm for calculating its values in the search domain, i.e.
+This present paper considers parallel algorithms for the solution of problems of multiextremal optimization.
+In multiextremal problems the possibility of a reliable estimation of the global optimum is essentially based on the availability of a priori information about the function, allowing us to link possible values of an optimizable function with known values in search points.
+Very often, such information about the solvable problem is presented as the assumption that the objective function $\varphi$ satisfies a Lipschitz condition with a priori unknown constant $L$.
+The objective function can be represented by the algorithm of calculation of its values in the search domain, i.e.
 it can be a ``black-box'' function.
-Similar problems are common in applications (problems of optimal design of objects and technological processes in various fields of technology, problems of model fitting according to observed data in scientific research, optimization of control processes, etc., see the review \cite{RefPinter}).
+Similar problems are widespread in applications (problems of optimal design of objects and technological processes in various technical domains, problems of model fitting according to observed data in scientific research, optimization of control processes, etc., see the review \cite{RefPinter}).
 
-Multiextremal optimization problems offer little in terms of analytical solution, while numerical solution is computationally challenging as computational costs often grow exponentially in dimension.
-By employing modern parallel computational systems it is possible to broaden the area of applicability of global optimization methods.
-At the same time, the problem of efficient parallelization of the global search process arises.
-Therefore, development of efficient parallel methods for numerical solution of multiextremal optimization problems and their software implementation for modern computational systems is the obvious focus of research in this area.
+Multiextremal optimization problems offer little in terms of analytical solution and numerical solution is computationally challenging as computational costs often grow exponentially in dimension.
+Employing modern parallel computational systems broadens the area of applicability of global optimization methods and, at the same time, poses a problem of efficient parallelization of the global search process.
+Therefore development of efficient parallel methods for numerical solution of multiextremal optimization problems and their software implementation for modern computational systems is the main direction of research in this area.
 
-One of the promising trends in the field of parallel global optimization (as in many areas connected with software implementation of complex algorithms) is the use of graphics processing units (GPUs).
+One of the promising directions of development in the domain of parallel global optimization (as in many areas connected with software implementation of complex algorithms) is the use of graphics processing units (GPUs).
 During the last decade, productivity of GPUs has been rapidly growing to satisfy the continuously growing demand of graphic application developers.
-Furthermore, some principles of development of graphics hardware have changed over the past few years.
+Furthermore, some principles of development of graphics hardware have changed over the past few years, and has become more programmable.
 Today a GPU is a high-performance flexibly programmable and massively parallel processor that can provide solution to many computationally complex problems \cite{RefHwu}.
 
-However, in the field of global optimization the potential of GPUs is not completely utilized.
-GPUs are generally used for parallelization of algorithms, which one way or another are based on the concept of random search (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi}); a review of this trend of GPU optimization algorithms is given in \cite{RefLangdon}.
+However, in the domain of global optimization the potential of GPUs is not completely unlocked.
+GPUs are generally used for parallelization of algorithms, which one way or another are based on the concept of random search (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi}); a review of this direction of GPU optimization algorithms is given in \cite{RefLangdon}.
 Owing to their stochastic nature, algorithms of this type generally do not guarantee convergence to a globally optimal solution, and that sets them apart from deterministic methods.
 
-Some parallel variants are proposed for many deterministic algorithms of lipschitzian global optimization with guaranteed convergence \cite{RefGergel2005,RefEvtushenko,RefHe,RefPaulavicius}.
+For many deterministic algorithms of Lipschitzian global optimization with guaranteed convergence parallel variants are proposed \cite{RefGergel2005,RefEvtushenko,RefHe,RefPaulavicius}.
 However, these versions of algorithms are parallelized on CPU using MPI and/or OpenMP technologies; there are no implementations on GPU.
@@
 Section \ref{sec:2} contains a description of the basic algorithm of global search.
-Section \ref{sec:3} contains a scheme of its parallelization, and section \ref{sec:4} includes theoretical statements that characterize the speedup and non-redundancy of the parallel algorithm.
-In section \ref{sec:5}, ways of possible implementation of the parallel algorithm on GPU are discussed.
-Section \ref{sec:6} contains the results of numerical experiments; a comparison of the sequential algorithm and parallel CPU and GPU versions for solving a series of test multidimensional multiextremal problems is also made.
+Section \ref{sec:3} contains a scheme of its parallelization, and section \ref{sec:4} includes theoretical statements that characterize speedup and non-redundancy of the parallel algorithm.
+In section \ref{sec:5}, ways of possible implementation of parallel algorithm on GPU are discussed.
+Section \ref{sec:6} contains the results of numerical experiments; comparison of the sequential algorithm and parallel CPU and GPU versions for the solution of a series of test multidimensional multiextremal problems is also made.
 Section \ref{sec:7} concludes the paper.
@@
 
-Let us consider the problem of search for a global minimum of an $N$-dimensional function $\varphi(y)$ within a hyper interval $D$
+Let us consider the problem of search for a global minimum of an $N$-dimensional function $\varphi(y)$ within hyper interval $D$
 \begin{eqnarray}\label{eq:1}
@@
 
-Let us assume that the function $\varphi$ satisfies a Lipschitz condition with an a priori unknown constant $L$
+Let us assume that function $\varphi$ satisfies a Lipschitz condition with a priori unknown constant $L$
 \[
@@
 
-There is a number of ways to adapt effective one-dimensional algorithms for solving multidimensional problems; see, for example, the diagonal partitions method in \cite{RefSergeyev2006} or the simplicial partitions method in \cite{RefZilinskas}.
+There is a number of ways to adapt effective one-dimensional algorithms for solution of multidimensional problems; see, for example, diagonal partitions method in \cite{RefSergeyev2006} or simplicial partitions method in \cite{RefZilinskas}.
 In this paper, we will use the approach based on the idea of dimension reduction by means of a Peano curve $y(x)$, which continuously and unambiguously maps the unit interval [0,1] onto the $n$-dimensional cube
@@
 Problems of numerical construction of Peano-type space filling curves and the corresponding theory are considered in detail in \cite{RefStrongin2000}, \cite{RefSergeyev2013}.
-Here we will note that a numerically constructed curve (\textit{evolvent}) is an approximation to a theoretical Peano curve with an accuracy $2^{-m}$ , where $m$ is an evolvent construction parameter.
-It means that using an evolvent we can distinguish points in a multidimensional space, the coordinates of which differ by not less than $2^{-m}$; closer points are considered as coincident.
-Thereby an evolvent constructed with a fixed $m$ allows obtaining a solution of a problem with an accuracy of not more than $2^{-m}$ by coordinate.
+Here we will note that a numerically constructed curve is an approximation to a theoretical Peano curve with accuracy $2^{-m}$ , where $m$ is a curve construction parameter.
 
-By using this kind of mapping it is possible to reduce the multidimensional problem~(\ref{eq:1}) to a one-dimensional problem
+Use of mapping of this kind allows a reduction of a multidimensional problem~(\ref{eq:1}) to a one-dimensional problem
 \[
@@
 \]
-An important property of such mapping is preservation of boundedness of function relative differences  (see \cite{RefStrongin2000}): if the function $\varphi(y)$ in the domain $D$ satisfies the Lipschitz condition, then the function $\varphi(y(x))$ on the interval $[0,1]$ will satisfy a uniform H{\"o}lder condition
+An important property is preservation of boundedness of function relative differences  (see \cite{RefStrongin2000}): if function $\varphi(y)$ in domain $D$ satisfies Lipschitz condition, then function $\varphi(y(x))$ on interval $[0,1]$ will satisfy a uniform H{\"o}lder condition
 \[
@@
 \]
-where the H{\"o}lder constant $H$ is linked to the Lipschitz constant $L$ by the relation
+where H{\"o}lder constant $H$ is linked to Lipschitz constant $L$ by the relation
 \[
@@
 \]
-satisfying the H{\"o}lder condition.
+satisfying H{\"o}lder condition.
 
-The considered algorithm for solving this problem (here, according to \cite{RefStrongin2000}) involves constructing a sequence of points $x^k$, where the values of minimized function $z^k = f(x^k)$ are calculated.
-Let us call the process of calculating the function value (including the construction of an image $y^k=y(x^k)$) the ``trial'', and the pair $(x^k, z^k)$, the ``trial result''.
-The set of pairs $\left\{(x^k, z^k)\right\}, 1\leq k\leq n,$ makes up the search data collected using the method after carrying out $n$ steps.
+The considered algorithm of the solution of this problem (here, according to \cite{RefStrongin2000}) presupposes construction of sequence of points $x^k$, where values of minimized function $z^k = f(x^k)$ are calculated.
+Let us call the process of calculation of the value of function (including construction of an image $y^k=y(x^k)$) ``trial'', and pair $(x^k, z^k)$ --  the ``trial result''.
+Set of pairs $\left\{(x^k, z^k)\right\}, 1\leq k\leq n,$ make up the search data collected using the method after carrying out $n$ steps.
 The rules that determine the work of the global search algorithm are as follows.
 
-At the first iteration of the method the trial is carried out at an arbitrary internal point $x^1$ of the interval $[0,1]$.
+At the first iteration of the method the trial is carried out in arbitrary internal point $x^1$ of interval $[0,1]$.
 The point of trial at the next iteration $(k+1)$ is determined according to these rules.
@@
 \]
-which includes boundary points of the interval $[0,1]$ and the points of the previous trials, with subscripts in increasing order of coordinate values, i.e.
+which includes boundary points of interval $[0,1]$ and points of the previous trials, with subscripts in increasing order of coordinate values, i.e.
 \[
@@
 
-Step 3. Calculate a \textit{characteristic} for every interval $(x_{i-1}, x_i), \; 1\leq i \leq k+1,$   according to the following formulas
+Step 3. Calculate \textit{characteristic} for every interval $(x_{i-1}, x_i), \; 1\leq i \leq k+1,$   according to the following formulas
 \[
@@
 
-Step 4. Find interval $(x_{t-1},x_t)$ with the maximum characteristic
+Step 4. Find interval $(x_{t-1},x_t)$ with maximum characteristic
 \[
@@
 
-Step 5. Carry out a trial at the point $x^{k+1}\in(x_{t-1},x_t)$, calculated using the following formulas
+Step 5. Carry out a trial in point $x^{k+1}\in(x_{t-1},x_t)$, calculated using the following formulas
 \[
-x^{k+1} = \frac{x_t+x_{t-1}}{2}, \textrm{ if } t=1 \textrm{ or } t=k+1,
+x^{k+1} = \frac{x_t+x_{t-1}}{2}, t=1, t=k+1,
 \]
 \[
-x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N, \textrm{ if } 1<t<k+1.
+x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N, 1<t<k+1.
 \]
 
-The algorithm terminates if the condition $\Delta_t<\epsilon$ is satisfied; here $\epsilon>0$ is the preset accuracy. For estimation of the global solution, values
+The algorithm stops working if the condition $\Delta_t<\epsilon$ is satisfied; here $\epsilon>0$ the accuracy is preset. For estimation of the global solution values
 \[
@@
 
-Remark 1.
-The normalized characteristics $R(i)$ from (\ref{eq:14}) can be viewed as probabilities of locating a global minimum within the interval $(x_{i-1},x_i), 1\leq i \leq k+1$.
-This statement is proved in \cite{RefStrongin2000}.
-Thus, at each iteration a new trial point is selected inside the interval, which has the greatest probability of finding a global minimum.
-
-Remark 2.
 The proof of convergence of this algorithm is provided in \cite{RefStrongin2000}.
-The modifications taking into account existence of inequality constraints in the problem and the information about the objective function derivative are given in \cite{RefBarkalov,RefGergel1996,RefGergel1997}.
+The modifications considering existence of inequality constraints in the problem and information about objective function derivative are given in \cite{RefBarkalov,RefGergel1996,RefGergel1997}.
 
@@
 \end{eqnarray}
-and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
-The experiment used the method parameters $r=3$, $\epsilon=10^{-3}$, and the evolvent construction parameter $m=12$.
-Fig.~\ref{fig:1} shows the contour lines of a function generated according to (\ref{eq:19}) and the points of 1103 trials carried out using the method before the required accuracy was obtained.
+and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in interval $[-1,1]$.
+The experiment used method parameters $r=3$, $\epsilon=10^{-3}$, and evolvent construction parameter $m=12$.
+Fig.~\ref{fig:1} shows contour lines of one of functions generated according to (\ref{eq:19}) and points of 1103 trials carried out using the method before the required accuracy was obtained.
 \begin{figure}
@@
   \includegraphics[width=0.75\textwidth]{fig1.jpg} 
-  \caption{Minimization of a two-dimensional function by the sequential algorithm}
+  \caption{Minimization of two-dimensional function}
   \label{fig:1}       % Give a unique label
@@
 
-The first option is to parallelize calculation of the target function describing the optimized object. This method can provide some speedup, but is specific for each problem.
+First, it is possible to parallelize calculation of the target function describing the optimized object. This method can provide speedup, but is specific for each solved problem.
 
-The second option is to parallelize implementation of algorithm computing rules for selection of the next trial point.
-In this case the way of parallelization will depend on the class of algorithms.
-Moreover, these rules are often rather simple and it is inexpedient to parallelize them (overhead costs of the organization of parallelism will reduce possible speedup to zero).
+Secondly, it is possible to parallelize implementation of algorithm computing rules providing selection of the point of the next trial.
+In this case the way of parallelization will depend on the class of algorithms and, besides, these rules are often rather simple and it is inexpedient to parallelize them (overhead costs of the organization of parallelism will reduce possible speedup to zero).
 
-The third option is to  change the scheme of an algorithm for the purpose of carrying out several trials in parallel.
-This approach is most promising since it is characterized by efficiency (the part of computing process, in which the bulk of calculations is carried out, is parallelized) and generality (it is applicable for a wide class of characteristical algorithms of multiextremal optimization \cite{RefGrishagin1997}).
+Thirdly, it is possible to change the scheme of algorithm for the purpose of carrying out several trials in parallel.
+This approach is the most promising since it is characterized by efficiency (the part of computing process, in which the bulk of calculations is carried out, is parallelized) and generality (it is applicable for a wide class of characteristical algorithms of multiextremal optimization \cite{RefGrishagin1997}).
 
 Let us describe a change of rules of global search algorithm for the purpose of simultaneous (synchronous) carrying out in one iteration of method $p$ of trials.
-Let us make $k(n)$ the total number of trials, carried out after $n$ parallel iterations.
+Let us make $k(n)$ the total number of trials, carried out after $n$ of parallel iterations.
 
-Suppose $n\geq 1$  iterations of the method are carried out, in which trials were performed at $k=k(n)$ points $x^i,1\leq i \leq k$.
+Suppose $n\geq 1$  iterations of the method are carried out, in which trials were performed in $k=k(n)$ points $x^i,1\leq i \leq k$.
 Then points $x^{k+1},\dots,x^{k+p}$  of search trials of the next $n+1$ iteration are determined according to the rules.
 
-Steps 1--3 of the parallel algorithm fully repeat steps 1--3 of the sequential global search algorithm.
+Steps 1--3 of the parallel algorithm fully repeat steps 1--3 of the global search algorithm.
 
@@
 
-Step 5. Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), 1\leq j\leq p$, calculated using the formulas
+Step 5. Carry out new trials in points $x^{k+j}\in(x_{t_j-1},x_{t_j}), 1\leq j\leq p$, calculated using formulas
 \[
-x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2}, \textrm{ if } t_j=1 \textrm{ or } t_j=k+1,
+x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2}, t_j=1, t_j=k+1,
 \]
 \[
-x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - \mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N, \textrm{ if } 1<t_j<k+1.
+x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - \mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N, 1<t_j<k+1.
 \]
 
-The algorithm terminates if the condition $\Delta_{t_j}<\epsilon$ is satisfied at least for one number $t_j, 1 \leq j \leq p$ ; $\epsilon>0$ is the preset accuracy.
-For estimation of the global solution, values
+The algorithm stops working if the condition $\Delta_{t_j}<\epsilon$ is satisfied at least for one number $t_j, 1 \leq j \leq p$ ; $\epsilon>0$ is the preset accuracy.
+For estimation of the global solution values
 \[
@@
 
-The modifications of the parallel algorithm with the account of the information about the minimized function derivative are given in \cite{RefGergel1999}.
+The modifications of the parallel algorithm considering information about the minimized function derivative are given in \cite{RefGergel1999}.
 
 This method of organizing parallel computing has the following justification \cite{RefStrongin2000,RefGrishagin1997}.
-The characteristics of intervals (\ref{eq:14}) used in the global search algorithm can be considered as probability measures of the global minimum point location in these intervals.
+The characteristics of intervals (\ref{eq:14}) used in the algorithm of global search can be considered as probability measures of location of the global minimum point in these intervals.
 Inequalities (\ref{eq:21}) arrange intervals according to their characteristics, and trials are carried out in parallel in the first $p$ intervals with the largest probabilities.
 
-Let us use a parallel algorithm of global search for solving a problem from section \ref{sec:2} with the numbers of parallel trials $p=2$ and $p=4$.
+Let us use a parallel algorithm of global search for solution of a problem from section \ref{sec:2} with numbers of parallel trials $p=2$ and $p=4$.
 With $p=2$ the number of iterations was $569$, and the number of trials was $1138$.
 With $p=4$ the number of iterations was reduced to $270$, and $1080$ trials were carried out.
-Fig.~\ref{fig:2} shows contour lines of the function with points of the trials performed by the method with $p=2$.
-Points of redundant trials carried out using a parallel algorithm are marked in the figure by "+".
-The concept of redundancy is discussed in more detail in the following section.
+Contour lines of the function are shown in Fig.~\ref{fig:2} with points of the trials carried out by the method when using $p=4$.
+Comparison of Fig.~\ref{fig:1} and Fig.~\ref{fig:2} shows absence of redundant trial points.
 \begin{figure}
@@
   \includegraphics[width=0.75\textwidth]{fig2.jpg} 
-  \caption{Minimization of a two-dimensional function by the parallel algorithm}
+  \caption{Minimization of a two-dimensional function by a parallel algorithm}
   \label{fig:2}
@@
 Let us describe theoretical properties of a parallel algorithm, which characterize its speedup.
-One of the main indicators of efficiency of parallel algorithms (in any field, not only in global optimization) is a speedup in time
+One of the main indicators of efficiency of parallel algorithms (in any domain, not only in global optimization) is speedup in time
 \[
@@
 \]
-where $T(1)$ is the time required to solve a problem by a sequential algorithm, and $T(p)$ is the time for solving the same problem by a parallel algorithm in a system with $p$ computing elements.
-The characteristic of efficiency of a parallel algorithms (in relation to optimization algorithms) is also a speedup in iterations
+where $T(1)$ is the time spent for solution of a problem by a sequential algorithm, and $T(p)$ is the time of solution of the same problem by a parallel algorithm in a system with $p$ computing elements.
+The characteristic of efficiency of parallel algorithms (in relation to algorithms of optimization) is also speedup in number of iterations
 \begin{equation}\label{eq:26}
@@
 \end{equation}
-where $n(1)$ is the number of trials carried out using the sequential method, and $n(p)$ is the number of trials carried out using the parallel method with $p$ processors.
-This characteristic is especially important since for solving applied problems the time required to carry out a trial exceeds the time required to process its results.
+where $n(1)$ is the number of the trials carried out using the sequential method, and $n(p)$ is the number of the trials carried out using the parallel method with p processors.
+This characteristic is especially important since for solution of applied problems the time of carrying out of a trial exceeds the time of processing of its results.
 
-It is obvious that the number of trials $n(p)$ for sequential and parallel algorithms described in sections \ref{sec:2} and \ref{sec:3} will differ.
+It is obvious that number of trials $n(p)$ for sequential and parallel algorithms described in sections \ref{sec:2} and \ref{sec:3} will differ.
 Actually, the sequential algorithm of global search when selecting point $x^{k+1}$ of the next $k+1$ trial possesses complete information received at the previous $k$ iterations.
-Based on the same information, the parallel algorithm of global search selects $p$ points $x^{k+j}, 1\leq j \leq p$, rather than one at iteration $k+1$.
-It means that selection of point $x^{k+j}$ is carried out in absence of information on the trial results at points $x^{k+i}, 1\leq i<j$.
+The parallel algorithm of global search selects not one, but $p$ points $x^{k+j}, 1\leq j \leq p$,  at iteration $k+1$ based on the same information.
+It means that selection of point $x^{k+j}$ is carried out in absence of information on the trial results in points $x^{k+i}, 1\leq i<j$.
 Only the first point $x^{k+1}$ will coincide with the point selected by the sequential algorithm.
-Other trials points generally cannot coincide with the points generated by the sequential algorithm.
-The efficiency offered by the use of parallel processors can be reduced when such trials are performed.
+Points of the other trials generally can not coincide with the points generated by the sequential algorithm.
+Carrying out of such trials can reduce efficiency of use of parallel processors.
 Therefore, let us consider such trials as ``redundant'', and the value
@@
 \]
-will represent ``method redundancy''.
+as ``method redundancy''.
 
-Let us set a series of trials $\left\{x^k\right\}$ and $\left\{y^m\right\}$ generated correspondingly by sequential and parallel algorithms for solving the same problem with $\epsilon = 0$ in the stopping condition.
+Let us set the series of trials $\left\{x^k\right\}$ and $\left\{y^m\right\}$ generated correspondingly by sequential and parallel algorithms for solution of the same problem with $\epsilon = 0$ in the condition of stopping.
 The following theorem from \cite{RefStrongin2000} determines the number of computing elements $p$, which can be involved for non-redundant parallelization.
@@
 	\item
-	For the value $M$ from (\ref{eq:11}) the following inequality
+	For value $M$ from (\ref{eq:11}) the following inequality
 	\[
@@
 	\]
-	holds, where $H$ is the H{\"o}lder constant of the minimized function.
+	holds, where $H$ is H{\"o}lder constant of the minimized function.
 \end{enumerate}
@@
 
-The theorem corollary plays a special role in solving multidimensional problems reduced to one-dimensional problems by means of the Peano-like evolvent $y(x)$.
-The evolvent $y(x)$, which is an approximation to the Peano curve, has the effect of ``splitting'' of a global minimum point $y^\ast\in D$ to several preimages in the interval $[0,1]$.
+The theorem corollary plays a special role for solution of multidimensional problems reduced to one-dimensional problems by means of Peano-like evolvent $y(x)$.
+Evolvent $y(x)$, which is approximation to Peano curve, has the effect of ``splitting'' of a point of the global minimum $y^\ast\in D$ to several preimages in interval $[0,1]$.
 If function $\varphi(y)$ has the only global minimum in $D$, the ``reduced'' function $f(x)$ can have up to $2^N$ local extremum points close (by value) to a global extremum point (see \cite{RefStrongin2000}).
-By applying a parallel global search algorithm for minimization of a similar function it is possible to achive non-redundancy when using up to $2^N+1$ computing elements.
+In case of applying a parallel global search algorithm for minimization of a similar function it is possible receive non-redundancy when using up to $2^N+1$ computing elements.
 
-The parallel algorithm of global search has demonstrated a good speedup in implementation on CPU in systems with a small number of processors.
+The parallel algorithm of global search showed a good speedup in implementation on CPU in systems with a small number of processors.
 In \cite{RefGrishagin1997,RefStrongin2003} a speedup of algorithm close to linear was experimentally verified when carrying out 2-4 parallel trials.
-The results obtained with the parallel GSA for minimization of 100 two-dimensional multiextremal functions described by formula (\ref{eq:19}) are shown in table \ref{tab:1}.
-All experiments used method parameters $r=2.9$, and $\epsilon=10^{-3}$, and the global minimum was found with the required accuracy in all $100$ problems.
+The results of work of the parallel GSA for minimization of 100 two-dimensional multiextremal functions described by formula (\ref{eq:19}) are shown in table \ref{tab:1}.
+All experiments used method parameters $r=2.9$, and $\epsilon=10^{-3}$, and the global minimum was found with the demanded accuracy in all $100$ problems.
 The table shows the number of used processors $p$, the average number of trials $n(p)$ rounded to integers and average speedup in iterations $s(p)$ from (\ref{eq:26}).
@@
 
-What are the advantages of GPU over CPU when implementing global optimization algorithms?
+What advantages does GPU have in case of implementation of global optimization algorithms?
 First, CPU has only a small (up to $16$ in the advanced models) number of cores working independently at a high clock frequency.
@@
 GPU, on the contrary, works at low clock frequencies and uses hundreds of simpler computing elements.
-Second, a considerable part of a CPU crystal is occupied by a high-speed cache memory, while practically the whole of GPU consists of arithmetic and logic units.
+Second, a considerable share of a CPU crystal is occupied by a high-speed cache memory, while practically the whole of GPU consists of arithmetic and logic units.
 Therefore, GPUs are especially effective in problems, in which the number of arithmetic operations is great in comparison with memory operations.
 
-With respect to methods of global optimization, parallel computing of many values of the objective function simultaneously is an operation which can be efficiently implemented on GPU.
-In the function value calculation procedure implemented on GPU transfers of information from CPU to GPU will be minimal: it is necessary only to transfer the coordinates of trial points to GPU and to receive the values of the function in these points.
-The functions that determine the processing of trial results according to algorithm and require working with a large volume of collected search information can be efficiently implemented on CPU.
+With respect to methods of global optimization, the operation which can be implemented effectively on GPU is simultaneous parallel computing of many values of the objective function.
+Naturally, for this purpose the function value calculation procedure has to be implemented on GPU.
+Transfers of information from CPU to GPU will be minimal: it is necessary only to transfer to GPU coordinates of trial points and to receive values of the function in these points.
+The functions that define processing of results of trials according to algorithm and require work with a large volume of collected search information can be effectively implemented on CPU.
 
-A general scheme showing the organization of calculations using GPU is given in Fig.~\ref{fig:3}.
-According to this scheme, steps 1 -- 4 of a parallel global search algorithm are carried out on a CPU.
-The coordinates of $p$ points of trials calculated at step 4 of the algorithm are collected in an intermediate buffer, and then are transferred to the graphics processing unit.
-Calculation of function values at these points is carried out on a GPU, and then the results of trials (again via the intermediate buffer) are transferred to the CPU.
+A general scheme of organization of calculations using GPU is shown in Fig.~\ref{fig:3}.
+According to this scheme steps 1 -- 4 of a parallel global search algorithm are carried out on a CPU.
+The coordinates of $p$ points of trials calculated on step 4 of the algorithm are collected in an intermediate buffer, and then are transferred to the graphics processing unit.
+Calculation of function values in these points is carried out on GPU, and then the results of trials (again via an intermediate buffer) are transferred to CPU.
 
@@
 The use of several hundred of parallel computing elements (as in modern GPU) might not give a speedup of the search process by hundreds of times.
-In this case, the conditions of the theorem of non-redundant parallelization can be violated: the number of local extremums will be less than the number of computing cores.
+In this case, conditions of the theorem of non-redundant parallelization can be violated: the number of local extremums will be less than number of computing cores.
 Then (according to the theorem and its corollary) a parallel GSA will generate redundant trial points.
-Nevertheless, despite certain redundancy, the overall operating time of a GPU algorithm will be less than the operating time of a CPU algorithm.
+Nevertheless, despite certain redundancy, the general operating time of a GPU algorithm will be less than the operating time of a CPU algorithm.
 This has been verified by computing experiments, the results of which are given in section \ref{sec:6}.
@@
 
-It should be noted that widely known test problems from the field of multidimensional global optimization are characterized by small time required to calculate the objective function value.
+Keep in mind that widely known test problems from the domain of multidimensional global optimization are characterized by small time of objective function values calculation.
 Usually, such a calculation is nothing else than a summation of several (according to problem dimension) values of elementary functions.
-For example, the well known Rastrigin test function, the global minimum of which is located in $y^\ast = (0,\dots,0)$ with $\varphi(y^\ast)=0$, is set by the expression
+For example, the well known Rastrigin’s test function, the global minimum of which is located in $y^\ast = (0,\dots,0)$ with $\varphi(y^\ast)=0$, is set by the expression
 \begin{equation}\label{eq:33}
@@
 
-Therefore, for the purpose of imitation of the computational complexity inherent to applied problems of optimization \cite{RefBarkalov2013}, calculation of the objective function in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (series summation from 20 thousand elements).
+Therefore, for the purpose of imitation of the computational complexity inherent to applied problems of optimization \cite{RefBarkalov2013}, calculation of the objective function in all performed experiments was complicated by additional calculations without changing the type of function and arrangement of its minimums (series summation from 20 thousand elements).
 
-It is noteworthy that an important parameter for starting any GPU program (unlike a parallel program on CPU) is not only the number of threads $p$, but also the size of the \textit{thread block}.
+It is noteworthy that the important parameter for starting any GPU program (unlike a parallel program on CPU) is not only the number of threads $p$, but also the size of the \textit{thread block}.
 Separate threads on GPU are grouped in blocks of threads of identical size, thus each thread block is executed on a separate multiprocessor.
 Threads in a thread block can interact effectively using information in shared memory and synchronization.
-For the purpose of assessing the importance of the size of a thread block during the execution period of a GPU algorithm let us perform a preliminary series of experiments with the test function (\ref{eq:33}) with $N=5$.
-Let us vary the number $p$ of the used GPU threads and the size $pB$ of a thread block.
-All the other parameters of the method do not change ($r=2.3$, $\epsilon = 10^{-3}$, the evolvent construction parameter $m=10$).
+For the purpose of assessment of the importance of the size of a thread block during the period of execution of a GPU algorithm let us perform a preliminary series of experiments with test function (\ref{eq:33}) with $N=5$.
+Let us vary number $p$ of used GPU cores and size $pB$ of a thread block.
+All the other parameters of the method do not change ($r=2.3$, $\epsilon = 10^{-3}$, evolvent construction parameter $m=10$).
 Table \ref{tab:2} shows the GPU algorithm operating time depending on $p$ and $pB$.
@@
 
-As is seen from this table, there is a speedup of the algorithm in the case of an increase in the number of trials at one iteration up to the number of cores on GPU, with subsequent saturation.
-Owing to a weak information dependence the size of the thread block $pB$ does not have decisive effect, therefore let us use $pB=64$ in all subsequent experiments.
-When carrying out this set of experiments the load on GPU was 15 -- 85 \% (depending on the number of threads); the load was measured by GPUZ program.
+As is seen from this table, there is a speedup of algorithm in the case of an increase in the number of trials at one iteration up to the number of cores on GPU, with subsequent saturation.
+Owing to a weak information dependence the size of thread block $pB$ does not have determining influence, therefore let us use $pB=64$ in all subsequent experiments.
 
-After the parameters of the GPU algorithm have been determined let us make the main computing experiments on a series of multidimensional multiextremal problems.
-In work \cite{RefGaviano} a GKLS generator allowing generation of multiextremal optimization problems with known properties (number of local minima, size of their domains of attraction, global minimum point, etc.) is described.
+After fixing of parameters of the GPU let us make the main computing experiments on a series of multidimensional multiextremal problems.
+In work \cite{RefGaviano} a GKLS generator allowing generation of multiextremal optimization problems with known properties (number of local minimums, size of their domains of attraction, global minimum point, value of function, etc.) is described.
 
-The results of numerical comparison of the three sequential algorithms -- DIRECT \cite{RefJones}, DIRECT\textit{l} \cite{RefGablonsky} and global search algorithm (GSA) from section \ref{sec:2} -- are provided below (results of the work of the first two algorithms are given in paper \cite{RefGaviano}).
-Numerical comparison was carried out on function classes \textit{Simple} and \textit{Hard} of dimension 4 and 5 from \cite{RefGaviano} since solving a problems of dimension 2 and 3 demands a small number of iterations and it is impractical to use GPU for solving these problems.
-Global minimum $y^\ast$ was considered to be found, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e.
+The results of numerical comparison of three sequential algorithms -- DIRECT \cite{RefJones}, DIRECT\textit{l} \cite{RefGablonsky} and global search algorithm (GSA) from section \ref{sec:2} -- are provided below (results of work of the first two algorithms are given in paper \cite{RefGaviano}).
+Numerical comparison was carried out on function classes \textit{Simple} and \textit{Hard} of dimension 4 and 5 from \cite{RefGaviano} since the solution of problems of dimension 2 and 3 demands a small number of iterations and use of GPU for solution of these problems is impractical.
+Global minimum $y^\ast$ was considered as found, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e.
 $\left\|y^k-y^\ast\right\|\leq\delta$.
 The size of the vicinity was selected (according to \cite{RefGaviano}) as $\delta = \left\|b-a\right\|\sqrt[N]{\Delta}$, $N$ -- problem dimension, $a$ and $b$ -- borders of search domain $D$, parameter $\Delta=10^{-6}$ at $N=4$ and $\Delta=10^{-7}$ at $N=5$.
-When using the GSA for the class \textit{Simple} parameter $r=4.5$ was selected, for the class \textit{Hard} $r=5.6$ was selected; evolvent construction parameter was fixed as $m=10$.
+When using the GSA method for class \textit{Simple} parameter $r=4.5$ was selected, for class \textit{Hard} $r=5.6$ was selected; Peano curve construction parameter was fixed as $m=10$.
 The maximum allowable number of iterations was $K_{max} = 10^6$.
 
-The average number of iterations $k_{av}$ performed by the method for solving a series of problems from these classes is shown in table \ref{tab:3}.
-Symbol ''$>$''  reflects a situation in which not all problems of the class were solved by the method.
+The average number of iterations $k_{av}$ performed by the method for solution of a series of problems from these classes is shown in table \ref{tab:3}.
+Symbol $>$ reflects a situation in which not all problems of a class were solved by a method.
 This means that the algorithm was stopped as the maximum allowable number of iterations $K_{max}$ was achieved.
-In this case, value $K_{max}=10^6$ was used for calculating the average value of the number of iterations $k_{av}$ that corresponds to the lower estimate of this average value.
+In this case, value $K_{max}=10^6$ was used for calculation of the average value of number of iterations $k_{av}$ that corresponds to the lower estimate of this average value.
 The number of unsolved problems is specified in brackets.
@@
 			5	& \textit{Simple} &	$>$16057(1) &	16758 &	15920 \\
-				& \textit{Hard} &	$>$217215(16) &	$>$269064(4) & $>$148342(4) \\
+				& \textit{Hard} &	$>$217215(16) &	$>$269064(4) & $>$148342 (4) \\
 		\noalign{\smallskip}\hline
@@
 
-Table \ref{tab:3} shows that GSA outperforms DIRECT and DIRECT\textit{l} methods on all classes of problems in terms of average number of iterations.
-In class 5-\textit{Hard} all the three methods failed to solve all problems: DIRECT did not solve 16 problems, DIRECT\textit{l} and GSA, 4 problems.
+Table \ref{tab:3} shows that GSA outperforms DIRECT and DIRECT\textit{l} methods on all classes of problems by average number of iterations.
+And in class 5-\textit{Hard} each of the methods solved not all problems: DIRECT did not solve 16 problems, DIRECT\textit{l} and GSA -- 4 problems.
 
-Let us estimate now the speedup when using the parallel global search algorithm implemented on CPU depending on the number $p$ of the used cores.
-Table \ref{tab:4} shows the average number of iterations performed by the method when solving a series of problems, and table \ref{tab:5} shows the average time required to solve one problem.
-Tables \ref{tab:6} and \ref{tab:7} show speedup in time $S(p)$ and iterations $s(p)$, respectively; speedup of the parallel algorithm was measured in relation to the sequential one ($p=1$).
+Let us estimate now the speedup when using a parallel global search algorithm implemented on CPU depending on number $p$ of used cores.
+Table \ref{tab:4} shows the average number of iterations performed by the method during solution of a series of problems, and table \ref{tab:5} shows the average time of solution of one problem.
+Tables \ref{tab:6} and \ref{tab:7} show speedup on time $S(p)$ and iterations $s(p)$ respectively; speedup of parallel algorithm was measured in relation to the sequential one ($p=1$).
 
@@
 
-The data show five-fold time speedup, when using 32 cores for all classes (except 5-\textit{Hard}).
-The lack of expected speedup in class 5-\textit{Hard} in transition from 16 to 32 cores is explained by an increase in the average number of iterations and by the fact that out of 32 cores 16 are physical, and other 16 are virtual in hyper-threading mode, i.e.
-a real gain in performance in transition from 16 to 32 cores is less than twofold.
+The data show quintuple time speedup, when using 32 cores for all classes (except 5-\textit{Hard}).
+Delay in class 5-\textit{Hard} in transition from 16 to 32 cores is explained by an increase in average number of iterations and by the fact that out of 32 cores 16 are physical, and other 16 are virtual in hyper-threading mode, i.e.
+a real gain in performance in transition from 16 to 32 cores is less than double.
 
 Let us make a similar experiment using a GPU version of the algorithm.
-Also let us fix the average number of iterations and average time required to solve one problem (see tables \ref{tab:last4} -- \ref{tab:last}).
-Let us measure the speedup and redundancy of GPU algorithm in relation to the CPU algorithm started with 32 cores (see column $p=32$ in tables \ref{tab:4} and \ref{tab:5}).
-The number of the used threads $p$ on GPU varied, all other parameters of the method did not change.
-The number of threads, which is greater, than the number of cores, was chosen to show the following effect: the number of iterations executed by the algorithm will be reduced in case of an increase of the threads number, while the algorithm operation time will not be reduced.
-The minimum time of operation is reached, if the number of used threads does not surpass the number of GPU cores.
+Also let us fix the average number of iterations and average time spent on one problem (see the following two tables).
+Let us measure the speedup and redundancy of GPU algorithm in relation to the CPU algorithm started with 32 cores (see column $p=32$ in tables 4 and 5).
+The number of the used $p$ GPU cores varied, all other parameters of the method did not change.
 
@@
 
-The results of our experiments show some obvious advantages of the GPU algorithm in comparison with the CPU algorithm.
+The results of experiments show advantage of the GPU algorithm in comparison with the CPU algorithm.
 The algorithm using the maximum quantity of cores of CPU works slower than any of the given GPU implementations.
 In addition, the results show a considerable speedup in the number of iterations.
-For example, on average only 353 parallel iterations on GPU were sufficient to solve problems of the most difficult class considered -- 5-\textit{Hard}, whereas even with the use of all computing CPU cores over eleven thousand iterations were required.
+For example, solution of problems of the most difficult of the classes considered -- 5-\textit{Hard} -- required on average only 353 parallel iterations on GPU, whereas even when using all computing CPU cores the number of iterations made was over eleven thousand.
 Finally, the GPU algorithm is almost non-redundant in relation to the CPU algorithm.
@@
 
-By now, parallel implementations have been suggested for the most of the existing sequential algorithms.
-A review of modern methods of parallel global optimization can be found, for example, in \cite{RefApuzzo}.
-The methods described in \cite{RefApuzzo} can be divided into two classes: deterministic methods of global optimization, and heuristic methods, which are one way or another based on the random search concept.
-Algorithms belonging to the latter class allow use of massive parallelism and are easily implemented on GPU (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi} and the review \cite{RefLangdon}).
+Summing up the results, let us note that use of graphics processing units for solution of global optimization problems is a new, promising direction of research as high efficiency of modern supercomputers is reached (generally) due to use of graphics accelerators.
 
-Implementations on GPU for deterministic algorithms of global optimization has not been suggested yet.
-For example, parallelization of algorithms based on the ideas of the branch and bound method is carried out for a small number of processors using OpenMP and MPI \cite{RefPaulavicius}.
-It is noteworthy that the branch and bound method is applied in other areas of optimization; in particular, combinatorial optimization and integer programming.
-Using GPU shows good results in these areas (see \cite{RefBoukedjar}, \cite{RefCarneiro}).
-It is explained by the following factor.
-Problems of this class (such as the Travelling Salesman Problem or the Knapsack problem) are characterized by short time of criterion value calculation in one point and, at the same time, relatively long time required to process the results.
-Therefore, the most labor-intensive operations here are branching and bound computation, which are implemented on GPU (see \cite{RefBoukedjar}, \cite{RefCarneiro}).
+This paper considers a parallel algorithm of solution of multidimensional multiextremal optimization problems and its implementation on GPU.
+This algorithm was developed within the framework of information-statistical approach.
+Theoretical information on properties of parallel algorithm characterizing conditions of non-redundant parallelization are provided.
 
-This paper considers the problems of lipschitzian global optimization that are frequently encountered in applications.
-They are characterized by a considerable time required to calculate the function value in one point (in comparison with the processing of the results of such calculation).
-Therefore, calculation of the value of an optimized function in one point is done using a GPU, and the main objective of the algorithm (working at CPU) is to efficiently select points of parallel trials.
-This is carried out in the parallel implementation of GSA on GPU.
-
-Summing up the results, let us note that use of graphics processing units for solving global optimization problems is a new, promising line of research as high efficiency of modern supercomputers is reached (generally) due to the use of graphics accelerators.
-
-This paper considers a parallel algorithm for solving multidimensional multiextremal optimization problems and its implementation on GPU.
-This algorithm was developed within the framework of information-statistical approach.
-Theoretical information on the properties of the parallel algorithm characterizing the conditions of non-redundant parallelization are provided.
-
-For the purpose of experimental demonstration of theoretical properties of the considered parallel algorithms, computing experiments on a series of several hundred test problems of different dimensions were performed.
-The results obtained with the sequential global search algorithm show its superiority over other known methods of global optimization.
+For the purpose of experimental demonstration of theoretical properties of the considered parallel algorithms, computing experiments on a series of several hundred test problems of different dimension were performed.
+The results of work of the sequential global search algorithm show its superiority over other known methods of global optimization.
 At the same time the parallel algorithm shows a good speedup and low redundancy (both GPU and CPU version), and the GPU version of the parallel algorithm shows a speedup of 2 to 6 times in comparison with the fastest parallel CPU version.
@@
 
-\bibitem{RefApuzzo}
-D'Apuzzo, M.,  Marino, M., Migdalas, A., Pardalos, P.M., Toraldo, G.: Parallel computing in global optimization.
-In Handbook of Parallel Computing and Statistics, Chapman \& Hall, London, 225--258 (2006)
-
-\bibitem{RefBoukedjar}
-Boukedjar A., Lalami M. E., El Baz D.: Parallel Branch and Bound on a CPU-GPU System. In 20th Int. Conf. on Parallel,  Distributed and Network-Based Processing, 392--398 (2012)
-
-\bibitem{RefCarneiro}
-Carneiro, T. et al.: A new parallel schema for branch-and-bound algorithms using GPGPU. In 23rd Int. Symposium on Computer Architecture and High Performance Computing, 41--47 (2011)
-
 \end{thebibliography}

### 2019-10-07 2016/JoGO_GPU/BarkalovGergel.tex
JOGO2016 paper next version

@@
 it can be a ``black-box'' function.
-Similar problems are widespread in applications (problems of optimal design of objects and technological processes in various technical domains, problems of model fitting according to observed data in scientific research, optimization of control processes, etc., see the review \cite{RefPinter}).
+Similar problems are widespread in applications (problems of optimal design of objects and technological processes in various fields of technology, problems of model fitting according to observed data in scientific research, optimization of control processes, etc., see the review \cite{RefPinter}).
 
@@
 
-One of the promising directions of development in the domain of parallel global optimization (as in many areas connected with software implementation of complex algorithms) is the use of graphics processing units (GPUs).
+One of the promising directions of development in the field of parallel global optimization (as in many areas connected with software implementation of complex algorithms) is the use of graphics processing units (GPUs).
 During the last decade, productivity of GPUs has been rapidly growing to satisfy the continuously growing demand of graphic application developers.
-Furthermore, some principles of development of graphics hardware have changed over the past few years, and has become more programmable.
+Furthermore, some principles of development of graphics hardware have changed over the past few years, and it became more programmable.
 Today a GPU is a high-performance flexibly programmable and massively parallel processor that can provide solution to many computationally complex problems \cite{RefHwu}.
@@
 Section \ref{sec:2} contains a description of the basic algorithm of global search.
-Section \ref{sec:3} contains a scheme of its parallelization, and section \ref{sec:4} includes theoretical statements that characterize speedup and non-redundancy of the parallel algorithm.
-In section \ref{sec:5}, ways of possible implementation of parallel algorithm on GPU are discussed.
-Section \ref{sec:6} contains the results of numerical experiments; comparison of the sequential algorithm and parallel CPU and GPU versions for the solution of a series of test multidimensional multiextremal problems is also made.
+Section \ref{sec:3} contains a scheme of its parallelization, and section \ref{sec:4} includes theoretical statements that characterize the speedup and non-redundancy of the parallel algorithm.
+In section \ref{sec:5}, ways of possible implementation of the parallel algorithm on GPU are discussed.
+Section \ref{sec:6} contains the results of numerical experiments; a comparison of the sequential algorithm and parallel CPU and GPU versions for solving a series of test multidimensional multiextremal problems is also made.
 Section \ref{sec:7} concludes the paper.
@@
 
-Let us consider the problem of search for a global minimum of an $N$-dimensional function $\varphi(y)$ within hyper interval $D$
+Let us consider the problem of search for a global minimum of an $N$-dimensional function $\varphi(y)$ within a hyper interval $D$
 \begin{eqnarray}\label{eq:1}
@@
 
-Let us assume that function $\varphi$ satisfies a Lipschitz condition with a priori unknown constant $L$
+Let us assume that the function $\varphi$ satisfies a Lipschitz condition with a priori unknown constant $L$
 \[
@@
 
-There is a number of ways to adapt effective one-dimensional algorithms for solution of multidimensional problems; see, for example, diagonal partitions method in \cite{RefSergeyev2006} or simplicial partitions method in \cite{RefZilinskas}.
+There is a number of ways to adapt effective one-dimensional algorithms for solving multidimensional problems; see, for example, the diagonal partitions method in \cite{RefSergeyev2006} or the simplicial partitions method in \cite{RefZilinskas}.
 In this paper, we will use the approach based on the idea of dimension reduction by means of a Peano curve $y(x)$, which continuously and unambiguously maps the unit interval [0,1] onto the $n$-dimensional cube
@@
 Problems of numerical construction of Peano-type space filling curves and the corresponding theory are considered in detail in \cite{RefStrongin2000}, \cite{RefSergeyev2013}.
-Here we will note that a numerically constructed curve is an approximation to a theoretical Peano curve with accuracy $2^{-m}$ , where $m$ is a curve construction parameter.
+Here we will note that a numerically constructed curve is an approximation to a theoretical Peano curve with an accuracy $2^{-m}$ , where $m$ is a curve construction parameter.
 
-Use of mapping of this kind allows a reduction of a multidimensional problem~(\ref{eq:1}) to a one-dimensional problem
+By using mapping of this kind it is possible to reduce the multidimensional problem~(\ref{eq:1}) to a one-dimensional problem
 \[
@@
 \]
-An important property is preservation of boundedness of function relative differences  (see \cite{RefStrongin2000}): if function $\varphi(y)$ in domain $D$ satisfies Lipschitz condition, then function $\varphi(y(x))$ on interval $[0,1]$ will satisfy a uniform H{\"o}lder condition
+An important property is preservation of boundedness of function relative differences  (see \cite{RefStrongin2000}): if the function $\varphi(y)$ in the domain $D$ satisfies the Lipschitz condition, then the function $\varphi(y(x))$ on the interval $[0,1]$ will satisfy a uniform H{\"o}lder condition
 \[
@@
 \]
-where H{\"o}lder constant $H$ is linked to Lipschitz constant $L$ by the relation
+where the H{\"o}lder constant $H$ is linked to the Lipschitz constant $L$ by the relation
 \[
@@
 \]
-satisfying H{\"o}lder condition.
+satisfying the H{\"o}lder condition.
 
-The considered algorithm of the solution of this problem (here, according to \cite{RefStrongin2000}) presupposes construction of sequence of points $x^k$, where values of minimized function $z^k = f(x^k)$ are calculated.
-Let us call the process of calculation of the value of function (including construction of an image $y^k=y(x^k)$) ``trial'', and pair $(x^k, z^k)$ --  the ``trial result''.
+The considered algorithm for solving this problem (here, according to \cite{RefStrongin2000}) assumes constructing a sequence of points $x^k$, where the values of minimized function $z^k = f(x^k)$ are calculated.
+Let us call the process of calculating the function value (including the construction of an image $y^k=y(x^k)$) the ``trial'', and the pair $(x^k, z^k)$ --  the ``trial result''.
 Set of pairs $\left\{(x^k, z^k)\right\}, 1\leq k\leq n,$ make up the search data collected using the method after carrying out $n$ steps.
@@
 \[
-x^{k+1} = \frac{x_t+x_{t-1}}{2}, t=1, t=k+1,
+x^{k+1} = \frac{x_t+x_{t-1}}{2}, \textrm{ if } t=1 \textrm{ or } t=k+1,
 \]
 \[
-x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N, 1<t<k+1.
+x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N, \textrm{ if } 1<t<k+1.
 \]
@@
 \[
-x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2}, t_j=1, t_j=k+1,
+x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2}, \textrm{ if } t_j=1 \textrm{ or } t_j=k+1,
 \]
 \[
-x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - \mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N, 1<t_j<k+1.
+x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - \mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N, \textrm{ if } 1<t_j<k+1.
 \]
@@
 Transfers of information from CPU to GPU will be minimal: it is necessary only to transfer to GPU coordinates of trial points and to receive values of the function in these points.
-The functions that define processing of results of trials according to algorithm and require work with a large volume of collected search information can be effectively implemented on CPU.
+The functions that determine the processing of trial results according to algorithm and require work with a large volume of collected search information can be effectively implemented on CPU.
 
@@
 According to this scheme steps 1 -- 4 of a parallel global search algorithm are carried out on a CPU.
-The coordinates of $p$ points of trials calculated on step 4 of the algorithm are collected in an intermediate buffer, and then are transferred to the graphics processing unit.
-Calculation of function values in these points is carried out on GPU, and then the results of trials (again via an intermediate buffer) are transferred to CPU.
+The coordinates of $p$ points of trials calculated at step 4 of the algorithm are collected in an intermediate buffer, and then are transferred to the graphics processing unit.
+Calculation of function values in these points is carried out on GPU, and then the results of trials (again via the intermediate buffer) are transferred to CPU.
 
@@
 The use of several hundred of parallel computing elements (as in modern GPU) might not give a speedup of the search process by hundreds of times.
-In this case, conditions of the theorem of non-redundant parallelization can be violated: the number of local extremums will be less than number of computing cores.
+In this case, conditions of the theorem of non-redundant parallelization can be violated: the number of local extremums will be less than the number of computing cores.
 Then (according to the theorem and its corollary) a parallel GSA will generate redundant trial points.
@@
 
-It is noteworthy that the important parameter for starting any GPU program (unlike a parallel program on CPU) is not only the number of threads $p$, but also the size of the \textit{thread block}.
+It is noteworthy that an important parameter for starting any GPU program (unlike a parallel program on CPU) is not only the number of threads $p$, but also the size of the \textit{thread block}.
 Separate threads on GPU are grouped in blocks of threads of identical size, thus each thread block is executed on a separate multiprocessor.
 Threads in a thread block can interact effectively using information in shared memory and synchronization.
-For the purpose of assessment of the importance of the size of a thread block during the period of execution of a GPU algorithm let us perform a preliminary series of experiments with test function (\ref{eq:33}) with $N=5$.
-Let us vary number $p$ of used GPU cores and size $pB$ of a thread block.
-All the other parameters of the method do not change ($r=2.3$, $\epsilon = 10^{-3}$, evolvent construction parameter $m=10$).
+For the purpose of assessing the importance of the size of a thread block during the period of execution of a GPU algorithm let us perform a preliminary series of experiments with the test function (\ref{eq:33}) with $N=5$.
+Let us vary the number $p$ of the used GPU cores and the size $pB$ of a thread block.
+All the other parameters of the method do not change ($r=2.3$, $\epsilon = 10^{-3}$, the evolvent construction parameter $m=10$).
 Table \ref{tab:2} shows the GPU algorithm operating time depending on $p$ and $pB$.
@@
 
-As is seen from this table, there is a speedup of algorithm in the case of an increase in the number of trials at one iteration up to the number of cores on GPU, with subsequent saturation.
-Owing to a weak information dependence the size of thread block $pB$ does not have determining influence, therefore let us use $pB=64$ in all subsequent experiments.
+As is seen from this table, there is a speedup of the algorithm in the case of an increase in the number of trials at one iteration up to the number of cores on GPU, with subsequent saturation.
+Owing to a weak information dependence the size of the thread block $pB$ does not have determining influence, therefore let us use $pB=64$ in all subsequent experiments.
 
-After fixing of parameters of the GPU let us make the main computing experiments on a series of multidimensional multiextremal problems.
-In work \cite{RefGaviano} a GKLS generator allowing generation of multiextremal optimization problems with known properties (number of local minimums, size of their domains of attraction, global minimum point, value of function, etc.) is described.
+After fixing the GPU parameters let us make the main computing experiments on a series of multidimensional multiextremal problems.
+In work \cite{RefGaviano} a GKLS generator allowing generation of multiextremal optimization problems with known properties (number of local minima, size of their domains of attraction, global minimum point, function value, etc.) is described.
 
-The results of numerical comparison of three sequential algorithms -- DIRECT \cite{RefJones}, DIRECT\textit{l} \cite{RefGablonsky} and global search algorithm (GSA) from section \ref{sec:2} -- are provided below (results of work of the first two algorithms are given in paper \cite{RefGaviano}).
-Numerical comparison was carried out on function classes \textit{Simple} and \textit{Hard} of dimension 4 and 5 from \cite{RefGaviano} since the solution of problems of dimension 2 and 3 demands a small number of iterations and use of GPU for solution of these problems is impractical.
-Global minimum $y^\ast$ was considered as found, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e.
+The results of numerical comparison of the three sequential algorithms -- DIRECT \cite{RefJones}, DIRECT\textit{l} \cite{RefGablonsky} and global search algorithm (GSA) from section \ref{sec:2} -- are provided below (results of work of the first two algorithms are given in paper \cite{RefGaviano}).
+Numerical comparison was carried out on function classes \textit{Simple} and \textit{Hard} of dimension 4 and 5 from \cite{RefGaviano} since the solution of problems of dimension 2 and 3 demands a small number of iterations and it is impractical to use GPU for solving these problems.
+Global minimum $y^\ast$ was considered to be found, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e.
 $\left\|y^k-y^\ast\right\|\leq\delta$.
 The size of the vicinity was selected (according to \cite{RefGaviano}) as $\delta = \left\|b-a\right\|\sqrt[N]{\Delta}$, $N$ -- problem dimension, $a$ and $b$ -- borders of search domain $D$, parameter $\Delta=10^{-6}$ at $N=4$ and $\Delta=10^{-7}$ at $N=5$.
-When using the GSA method for class \textit{Simple} parameter $r=4.5$ was selected, for class \textit{Hard} $r=5.6$ was selected; Peano curve construction parameter was fixed as $m=10$.
+When using the GSA for the class \textit{Simple} parameter $r=4.5$ was selected, for the class \textit{Hard} $r=5.6$ was selected; Peano curve construction parameter was fixed as $m=10$.
 The maximum allowable number of iterations was $K_{max} = 10^6$.
 
-The average number of iterations $k_{av}$ performed by the method for solution of a series of problems from these classes is shown in table \ref{tab:3}.
-Symbol $>$ reflects a situation in which not all problems of a class were solved by a method.
+The average number of iterations $k_{av}$ performed by the method for solving a series of problems from these classes is shown in table \ref{tab:3}.
+Symbol $>$ reflects a situation in which not all problems of the class were solved by the method.
 This means that the algorithm was stopped as the maximum allowable number of iterations $K_{max}$ was achieved.
-In this case, value $K_{max}=10^6$ was used for calculation of the average value of number of iterations $k_{av}$ that corresponds to the lower estimate of this average value.
+In this case, value $K_{max}=10^6$ was used for calculating the average value of the number of iterations $k_{av}$ that corresponds to the lower estimate of this average value.
 The number of unsolved problems is specified in brackets.
@@
 
-Let us estimate now the speedup when using a parallel global search algorithm implemented on CPU depending on number $p$ of used cores.
-Table \ref{tab:4} shows the average number of iterations performed by the method during solution of a series of problems, and table \ref{tab:5} shows the average time of solution of one problem.
-Tables \ref{tab:6} and \ref{tab:7} show speedup on time $S(p)$ and iterations $s(p)$ respectively; speedup of parallel algorithm was measured in relation to the sequential one ($p=1$).
+Let us estimate now the speedup when using the parallel global search algorithm implemented on CPU depending on the number $p$ of the used cores.
+Table \ref{tab:4} shows the average number of iterations performed by the method when solving a series of problems, and table \ref{tab:5} shows the average time requared to solve one problem.
+Tables \ref{tab:6} and \ref{tab:7} show speedup in time $S(p)$ and iterations $s(p)$ respectively; speedup of the parallel algorithm was measured in relation to the sequential one ($p=1$).
 

### 2019-10-07 2016/JoGO_GPU/BarkalovGergel.tex
JOGO2016 paper next version

@@
 This present paper considers parallel algorithms for the solution of problems of multiextremal optimization.
+The objective function in such problems is characterized by existence of several local optima (usually their number is not known and can be large).
 In multiextremal problems the possibility of a reliable estimation of the global optimum is essentially based on the availability of a priori information about the function, allowing us to link possible values of an optimizable function with known values in search points.
@@
 Problems of numerical construction of Peano-type space filling curves and the corresponding theory are considered in detail in \cite{RefStrongin2000}, \cite{RefSergeyev2013}.
-Here we will note that a numerically constructed curve is an approximation to a theoretical Peano curve with an accuracy $2^{-m}$ , where $m$ is a curve construction parameter.
+Here we will note that a numerically constructed curve (\textit{evolvent}) is an approximation to a theoretical Peano curve with an accuracy $2^{-m}$ , where $m$ is an evolvent construction parameter.
+It means that using evolvent we can distinguish points in a multidimensional space, the coordinates of which differ by not less than $2^{-m}$; closer points are considered as coincident.
+Thereby an evolvent constructed with a fixed m allows obtaining a solution of a problem with an accuracy of not more than $2^{-m}$ by coordinate.
 
@@
 
+Remark 1.
+The normalized characteristics $R(i)$ from (\ref{eq:14}) can be viewed as probabilities of locating of a global minimize within the interval $[x_{i-1},x_i], 1\leq i \leq k+1$.
+This statement is proved in \cite{RefStrongin2000}.
+Thus, at each iteration a new trial point is selected inside the interval, which has the greatest probability of finding a global minimum.
+
+Remark 2.
 The proof of convergence of this algorithm is provided in \cite{RefStrongin2000}.
@@
 Contour lines of the function are shown in Fig.~\ref{fig:2} with points of the trials carried out by the method when using $p=4$.
-Comparison of Fig.~\ref{fig:1} and Fig.~\ref{fig:2} shows absence of redundant trial points.
+Points of redundant trials carried out using a parallel algorithm are marked in the figure by "+".
+In more detail the concept of redundancy is discussed in the following section.
 \begin{figure}
@@
 For the purpose of assessing the importance of the size of a thread block during the period of execution of a GPU algorithm let us perform a preliminary series of experiments with the test function (\ref{eq:33}) with $N=5$.
-Let us vary the number $p$ of the used GPU cores and the size $pB$ of a thread block.
+Let us vary the number $p$ of the used GPU threads and the size $pB$ of a thread block.
 All the other parameters of the method do not change ($r=2.3$, $\epsilon = 10^{-3}$, the evolvent construction parameter $m=10$).
@@
 Owing to a weak information dependence the size of the thread block $pB$ does not have determining influence, therefore let us use $pB=64$ in all subsequent experiments.
+When carrying out this set of experiments the load on GPU was 15 -- 85 \% (depending on the number of threads); the load was measured by GPUZ program.
 
@@
 Let us make a similar experiment using a GPU version of the algorithm.
-Also let us fix the average number of iterations and average time spent on one problem (see the following two tables).
-Let us measure the speedup and redundancy of GPU algorithm in relation to the CPU algorithm started with 32 cores (see column $p=32$ in tables 4 and 5).
-The number of the used $p$ GPU cores varied, all other parameters of the method did not change.
+Also let us fix the average number of iterations and average time spent on one problem (see the tables \ref{tab:last4} -- \ref{tab:last}).
+Let us measure the speedup and redundancy of GPU algorithm in relation to the CPU algorithm started with 32 cores (see column $p=32$ in tables \ref{tab:4} and \ref{tab:5}).
+The number of the used threads $p$ on GPU varied, all other parameters of the method did not change.
+The number of threads, which is greater, than the number of cores, was chosen to show the following effect: the number of iterations executed by an algorithm will be reduced in case of an increase in number of threads, while time of operation will not.
+The minimum time of operation is reached, if the number of used threads does not surpass the number of GPU cores.
 
@@
 
+By now, parallel implementations of the majority of the developed sequential algorithms have been suggested.
+Please find a review of modern methods of parallel global optimization, for example, in \cite{RefApuzzo}.
+
+The methods described in \cite{RefApuzzo} can be divided into two classes: deterministic methods of global optimization (including the algorithm of global search considered in this article), and heuristic methods, which are one way or another based on the concept random search.
+Algorithms belonging to the latter class provide for use of massive parallelism and are easily implemented on GPU (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi} and the review \cite{RefLangdon}).
+
+Global optimization of implementations on GPU for deterministic algorithms has not been suggested yet.
+For example, parallelization of algorithms based on the ideas of the branch and bound method is carried out for a small number of processors using OpenMP and MPI \cite{RefPaulavicius}.
+It is noteworthy that the branch and bound method is applied in other sections of optimization; in particular, combinatorial optimization and integer programming.
+Using GPU shows good results in these domains (see \cite{RefBoukedjar}, \cite{RefCarneiro}).
+It is explained by the following factor.
+Problems of this class (such as travelling salesman problem or knapsack problem) are characterized by short time of criterion value calculation in one point and, at the same time, relatively long time of results processing.
+Therefore, the most labor-intensive operations here are branching and bound computation, which are implemented on GPU (see \cite{RefBoukedjar}, \cite{RefCarneiro}).
+
+This paper considers the problems of lipshitzian global optimization, that are frequently met in applications.
+They are characterized by a considerable time of calculation of function value in one point (in comparison with processing of results of such calculation).
+Therefore, calculation of the value of an optimized function in one point is done via GPU, and the main objective of the algorithm (working at CPU) is to efficiently select points of parallel trials.
+This is carried out in the parallel implementation of GSA on GPU.
+
 Summing up the results, let us note that use of graphics processing units for solution of global optimization problems is a new, promising direction of research as high efficiency of modern supercomputers is reached (generally) due to use of graphics accelerators.
@@
 
+\bibitem{RefApuzzo}
+M.
+D’Apuzzo, M.
+Marino, A.
+Migdalas, P.M.
+Pardalos, and G.
+Toraldo, Parallel computing in global optimization, in Handbook of Parallel Computing and Statistics, E.J.
+Kontoghiorghes, ed.
+Chapman \& Hall, London, 2006, pp.
+225–258
+
+\bibitem{RefBoukedjar}
+Boukedjar A., Lalami M. E., El Baz D. Parallel Branch and Bound on a CPU-GPU System //PDP. – 2012. – С. 392-398
+
+\bibitem{RefCarneiro}
+Carneiro T.
+et al.
+A new parallel schema for branch-and-bound algorithms using GPGPU //Computer Architecture and High Performance Computing (SBAC-PAD), 2011 23rd International Symposium on.
+– IEEE, 2011.
+– С.
+41-47.
+
 \end{thebibliography}

### 2019-10-07 2016/JoGO_GPU/BarkalovGergel.tex
JOGO2016 paper next version

@@
 \begin{abstract}
-This work considers a parallel algorithm of solution of multidimensional multiextremal optimization problems.
+This work considers a parallel algorithm for solving multidimensional multiextremal optimization problems.
 This algorithm uses Peano-type space filling curves for dimension reduction.
-Conditions of non-redundant parallelization of the algorithm is considered.
-The issue of implementation of the algorithm on modern computing systems with use of a graphics processing unit (GPU) is investigated.
+Conditions of non-redundant parallelization of the algorithm are considered.
+Implementation of the algorithm on modern computing systems with the use of a graphics processing unit (GPU) is investigated.
 Speedup of the algorithm using GPU as compared with the same algorithm implemented on CPU only is demonstrated experimentally.
-Computing experiments are carried out on a series of several hundred of multidimensional multiextremal problems.
+Computing experiments are carried out on a series of several hundred multidimensional multiextremal problems.
 \keywords{Global optimization \and Parallel computing \and Graphics processing unit \and Peano-type space filling curves \and Characteristical algorithms}
@@
 
-This present paper considers parallel algorithms for the solution of problems of multiextremal optimization.
+This present paper considers parallel algorithms for solving multiextremal optimization problems.
 The objective function in such problems is characterized by existence of several local optima (usually their number is not known and can be large).
-In multiextremal problems the possibility of a reliable estimation of the global optimum is essentially based on the availability of a priori information about the function, allowing us to link possible values of an optimizable function with known values in search points.
-Very often, such information about the solvable problem is presented as the assumption that the objective function $\varphi$ satisfies a Lipschitz condition with a priori unknown constant $L$.
-The objective function can be represented by the algorithm of calculation of its values in the search domain, i.e.
+In multiextremal problems the possibility of a reliable estimation of the global optimum is essentially based on the availability of an a priori information about the function, allowing us to link possible values of an optimizable function with known values at search points.
+Very often, such information about the problem is presented as the assumption that the objective function $\varphi$ satisfies a Lipschitz condition with an a priori unknown constant $L$.
+The objective function can be represented by the algorithm for calculating its values in the search domain, i.e.
 it can be a ``black-box'' function.
-Similar problems are widespread in applications (problems of optimal design of objects and technological processes in various fields of technology, problems of model fitting according to observed data in scientific research, optimization of control processes, etc., see the review \cite{RefPinter}).
+Similar problems are common in applications (problems of optimal design of objects and technological processes in various fields of technology, problems of model fitting according to observed data in scientific research, optimization of control processes, etc., see the review \cite{RefPinter}).
 
-Multiextremal optimization problems offer little in terms of analytical solution and numerical solution is computationally challenging as computational costs often grow exponentially in dimension.
-Employing modern parallel computational systems broadens the area of applicability of global optimization methods and, at the same time, poses a problem of efficient parallelization of the global search process.
-Therefore development of efficient parallel methods for numerical solution of multiextremal optimization problems and their software implementation for modern computational systems is the main direction of research in this area.
+Multiextremal optimization problems offer little in terms of analytical solution, while numerical solution is computationally challenging as computational costs often grow exponentially in dimension.
+By employing modern parallel computational systems it is possible to broaden the area of applicability of global optimization methods.
+At the same time, the problem of efficient parallelization of the global search process arises.
+Therefore, development of efficient parallel methods for numerical solution of multiextremal optimization problems and their software implementation for modern computational systems is the obvious focus of research in this area.
 
-One of the promising directions of development in the field of parallel global optimization (as in many areas connected with software implementation of complex algorithms) is the use of graphics processing units (GPUs).
+One of the promising trends in the field of parallel global optimization (as in many areas connected with software implementation of complex algorithms) is the use of graphics processing units (GPUs).
 During the last decade, productivity of GPUs has been rapidly growing to satisfy the continuously growing demand of graphic application developers.
-Furthermore, some principles of development of graphics hardware have changed over the past few years, and it became more programmable.
+Furthermore, some principles of development of graphics hardware have changed over the past few years.
+Now graphics hardware is more programmable.
 Today a GPU is a high-performance flexibly programmable and massively parallel processor that can provide solution to many computationally complex problems \cite{RefHwu}.
 
-However, in the domain of global optimization the potential of GPUs is not completely unlocked.
-GPUs are generally used for parallelization of algorithms, which one way or another are based on the concept of random search (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi}); a review of this direction of GPU optimization algorithms is given in \cite{RefLangdon}.
+However, in the field of global optimization the potential of GPUs is not completely utilized.
+GPUs are generally used for parallelization of algorithms, which one way or another are based on the concept of random search (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi}); a review of this trend of GPU optimization algorithms is given in \cite{RefLangdon}.
 Owing to their stochastic nature, algorithms of this type generally do not guarantee convergence to a globally optimal solution, and that sets them apart from deterministic methods.
 
-For many deterministic algorithms of Lipschitzian global optimization with guaranteed convergence parallel variants are proposed \cite{RefGergel2005,RefEvtushenko,RefHe,RefPaulavicius}.
+Some parallel variants are proposed for many deterministic algorithms of lipschitzian global optimization with guaranteed convergence \cite{RefGergel2005,RefEvtushenko,RefHe,RefPaulavicius}.
 However, these versions of algorithms are parallelized on CPU using MPI and/or OpenMP technologies; there are no implementations on GPU.
@@
 
-Let us assume that the function $\varphi$ satisfies a Lipschitz condition with a priori unknown constant $L$
+Let us assume that the function $\varphi$ satisfies a Lipschitz condition with an a priori unknown constant $L$
 \[
@@
 Here we will note that a numerically constructed curve (\textit{evolvent}) is an approximation to a theoretical Peano curve with an accuracy $2^{-m}$ , where $m$ is an evolvent construction parameter.
-It means that using evolvent we can distinguish points in a multidimensional space, the coordinates of which differ by not less than $2^{-m}$; closer points are considered as coincident.
+It means that using an evolvent we can distinguish points in a multidimensional space, the coordinates of which differ by not less than $2^{-m}$; closer points are considered as coincident.
 Thereby an evolvent constructed with a fixed m allows obtaining a solution of a problem with an accuracy of not more than $2^{-m}$ by coordinate.
 
-By using mapping of this kind it is possible to reduce the multidimensional problem~(\ref{eq:1}) to a one-dimensional problem
+By using this kind of mapping it is possible to reduce the multidimensional problem~(\ref{eq:1}) to a one-dimensional problem
 \[
@@
 \]
-An important property is preservation of boundedness of function relative differences  (see \cite{RefStrongin2000}): if the function $\varphi(y)$ in the domain $D$ satisfies the Lipschitz condition, then the function $\varphi(y(x))$ on the interval $[0,1]$ will satisfy a uniform H{\"o}lder condition
+An important property of such mapping is preservation of boundedness of function relative differences  (see \cite{RefStrongin2000}): if the function $\varphi(y)$ in the domain $D$ satisfies the Lipschitz condition, then the function $\varphi(y(x))$ on the interval $[0,1]$ will satisfy a uniform H{\"o}lder condition
 \[
@@
 
-The considered algorithm for solving this problem (here, according to \cite{RefStrongin2000}) assumes constructing a sequence of points $x^k$, where the values of minimized function $z^k = f(x^k)$ are calculated.
-Let us call the process of calculating the function value (including the construction of an image $y^k=y(x^k)$) the ``trial'', and the pair $(x^k, z^k)$ --  the ``trial result''.
-Set of pairs $\left\{(x^k, z^k)\right\}, 1\leq k\leq n,$ make up the search data collected using the method after carrying out $n$ steps.
+The considered algorithm for solving this problem (here, according to \cite{RefStrongin2000}) involves constructing a sequence of points $x^k$, where the values of minimized function $z^k = f(x^k)$ are calculated.
+Let us call the process of calculating the function value (including the construction of an image $y^k=y(x^k)$) the ``trial'', and the pair $(x^k, z^k)$, the ``trial result''.
+The set of pairs $\left\{(x^k, z^k)\right\}, 1\leq k\leq n,$ makes up the search data collected using the method after carrying out $n$ steps.
 The rules that determine the work of the global search algorithm are as follows.
 
-At the first iteration of the method the trial is carried out in arbitrary internal point $x^1$ of interval $[0,1]$.
+At the first iteration of the method the trial is carried out at an arbitrary internal point $x^1$ of the interval $[0,1]$.
 The point of trial at the next iteration $(k+1)$ is determined according to these rules.
@@
 \]
-which includes boundary points of interval $[0,1]$ and points of the previous trials, with subscripts in increasing order of coordinate values, i.e.
+which includes boundary points of the interval $[0,1]$ and the points of the previous trials, with subscripts in increasing order of coordinate values, i.e.
 \[
@@
 
-Step 3. Calculate \textit{characteristic} for every interval $(x_{i-1}, x_i), \; 1\leq i \leq k+1,$   according to the following formulas
+Step 3. Calculate a \textit{characteristic} for every interval $(x_{i-1}, x_i), \; 1\leq i \leq k+1,$   according to the following formulas
 \[
@@
 
-Step 4. Find interval $(x_{t-1},x_t)$ with maximum characteristic
+Step 4. Find interval $(x_{t-1},x_t)$ with the maximum characteristic
 \[
@@
 
-Step 5. Carry out a trial in point $x^{k+1}\in(x_{t-1},x_t)$, calculated using the following formulas
+Step 5. Carry out a trial at the point $x^{k+1}\in(x_{t-1},x_t)$, calculated using the following formulas
 \[
@@
 
-The algorithm stops working if the condition $\Delta_t<\epsilon$ is satisfied; here $\epsilon>0$ the accuracy is preset. For estimation of the global solution values
+The algorithm terminates if the condition $\Delta_t<\epsilon$ is satisfied; here $\epsilon>0$ the accuracy is preset. For estimation of the global solution, values
 \[
@@
 Remark 1.
-The normalized characteristics $R(i)$ from (\ref{eq:14}) can be viewed as probabilities of locating of a global minimize within the interval $[x_{i-1},x_i], 1\leq i \leq k+1$.
+The normalized characteristics $R(i)$ from (\ref{eq:14}) can be viewed as probabilities of locating a global minimum within the interval $[x_{i-1},x_i], 1\leq i \leq k+1$.
 This statement is proved in \cite{RefStrongin2000}.
@@
 The proof of convergence of this algorithm is provided in \cite{RefStrongin2000}.
-The modifications considering existence of inequality constraints in the problem and information about objective function derivative are given in \cite{RefBarkalov,RefGergel1996,RefGergel1997}.
+The modifications taking into account existence of inequality constraints in the problem and the information about the objective function derivative are given in \cite{RefBarkalov,RefGergel1996,RefGergel1997}.
 
@@
 \end{eqnarray}
-and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in interval $[-1,1]$.
-The experiment used method parameters $r=3$, $\epsilon=10^{-3}$, and evolvent construction parameter $m=12$.
-Fig.~\ref{fig:1} shows contour lines of one of functions generated according to (\ref{eq:19}) and points of 1103 trials carried out using the method before the required accuracy was obtained.
+and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
+The experiment used the method parameters $r=3$, $\epsilon=10^{-3}$, and the evolvent construction parameter $m=12$.
+Fig.~\ref{fig:1} shows the contour lines of a function generated according to (\ref{eq:19}) and the points of 1103 trials carried out using the method before the required accuracy was obtained.
 \begin{figure}
@@
   \includegraphics[width=0.75\textwidth]{fig1.jpg} 
-  \caption{Minimization of two-dimensional function}
+  \caption{Minimization of a two-dimensional function by the sequential algorithm}
   \label{fig:1}       % Give a unique label
@@
 
-First, it is possible to parallelize calculation of the target function describing the optimized object. This method can provide speedup, but is specific for each solved problem.
+The first option is to parallelize calculation of the target function describing the optimized object. This method can provide some speedup, but is specific for each problem.
 
-Secondly, it is possible to parallelize implementation of algorithm computing rules providing selection of the point of the next trial.
-In this case the way of parallelization will depend on the class of algorithms and, besides, these rules are often rather simple and it is inexpedient to parallelize them (overhead costs of the organization of parallelism will reduce possible speedup to zero).
+The second option is to parallelize implementation of algorithm computing rules for selection of the next trial point.
+In this case the way of parallelization will depend on the class of algorithms.
+Moreover, these rules are often rather simple and it is inexpedient to parallelize them (overhead costs of the organization of parallelism will reduce possible speedup to zero).
 
-Thirdly, it is possible to change the scheme of algorithm for the purpose of carrying out several trials in parallel.
-This approach is the most promising since it is characterized by efficiency (the part of computing process, in which the bulk of calculations is carried out, is parallelized) and generality (it is applicable for a wide class of characteristical algorithms of multiextremal optimization \cite{RefGrishagin1997}).
+The third option is to  change the scheme of an algorithm for the purpose of carrying out several trials in parallel.
+This approach is most promising since it is characterized by efficiency (the part of computing process, in which the bulk of calculations is carried out, is parallelized) and generality (it is applicable for a wide class of characteristical algorithms of multiextremal optimization \cite{RefGrishagin1997}).
 
 Let us describe a change of rules of global search algorithm for the purpose of simultaneous (synchronous) carrying out in one iteration of method $p$ of trials.
-Let us make $k(n)$ the total number of trials, carried out after $n$ of parallel iterations.
+Let us make $k(n)$ the total number of trials, carried out after $n$ parallel iterations.
 
-Suppose $n\geq 1$  iterations of the method are carried out, in which trials were performed in $k=k(n)$ points $x^i,1\leq i \leq k$.
+Suppose $n\geq 1$  iterations of the method are carried out, in which trials were performed at $k=k(n)$ points $x^i,1\leq i \leq k$.
 Then points $x^{k+1},\dots,x^{k+p}$  of search trials of the next $n+1$ iteration are determined according to the rules.
@@
 
-Step 5. Carry out new trials in points $x^{k+j}\in(x_{t_j-1},x_{t_j}), 1\leq j\leq p$, calculated using formulas
+Step 5. Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), 1\leq j\leq p$, calculated using the formulas
 \[
@@
 
-The algorithm stops working if the condition $\Delta_{t_j}<\epsilon$ is satisfied at least for one number $t_j, 1 \leq j \leq p$ ; $\epsilon>0$ is the preset accuracy.
-For estimation of the global solution values
+The algorithm terminates if the condition $\Delta_{t_j}<\epsilon$ is satisfied at least for one number $t_j, 1 \leq j \leq p$ ; $\epsilon>0$ is the preset accuracy.
+For estimation of the global solution, values
 \[
@@
 
-The modifications of the parallel algorithm considering information about the minimized function derivative are given in \cite{RefGergel1999}.
+The modifications of the parallel algorithm with the account of the information about the minimized function derivative are given in \cite{RefGergel1999}.
 
 This method of organizing parallel computing has the following justification \cite{RefStrongin2000,RefGrishagin1997}.
-The characteristics of intervals (\ref{eq:14}) used in the algorithm of global search can be considered as probability measures of location of the global minimum point in these intervals.
+The characteristics of intervals (\ref{eq:14}) used in the global search algorithm can be considered as probability measures of the global minimum point location in these intervals.
 Inequalities (\ref{eq:21}) arrange intervals according to their characteristics, and trials are carried out in parallel in the first $p$ intervals with the largest probabilities.
 
-Let us use a parallel algorithm of global search for solution of a problem from section \ref{sec:2} with numbers of parallel trials $p=2$ and $p=4$.
+Let us use a parallel algorithm of global search for solving a problem from section \ref{sec:2} with the numbers of parallel trials $p=2$ and $p=4$.
 With $p=2$ the number of iterations was $569$, and the number of trials was $1138$.
 With $p=4$ the number of iterations was reduced to $270$, and $1080$ trials were carried out.
-Contour lines of the function are shown in Fig.~\ref{fig:2} with points of the trials carried out by the method when using $p=4$.
+Fig.~\ref{fig:2} shows contour lines of the function with points of the trials performed by the method with $p=2$.
 Points of redundant trials carried out using a parallel algorithm are marked in the figure by "+".
-In more detail the concept of redundancy is discussed in the following section.
+The concept of redundancy is discussed in more detail in the following section.
 \begin{figure}
@@
   \includegraphics[width=0.75\textwidth]{fig2.jpg} 
-  \caption{Minimization of a two-dimensional function by a parallel algorithm}
+  \caption{Minimization of a two-dimensional function by the parallel algorithm}
   \label{fig:2}
@@
 Let us describe theoretical properties of a parallel algorithm, which characterize its speedup.
-One of the main indicators of efficiency of parallel algorithms (in any domain, not only in global optimization) is speedup in time
+One of the main indicators of efficiency of parallel algorithms (in any field, not only in global optimization) is a speedup in time
 \[
@@
 \]
-where $T(1)$ is the time spent for solution of a problem by a sequential algorithm, and $T(p)$ is the time of solution of the same problem by a parallel algorithm in a system with $p$ computing elements.
-The characteristic of efficiency of parallel algorithms (in relation to algorithms of optimization) is also speedup in number of iterations
+where $T(1)$ is the time required to solve a problem by a sequential algorithm, and $T(p)$ is the time for solving the same problem by a parallel algorithm in a system with $p$ computing elements.
+The characteristic of efficiency of a parallel algorithms (in relation to optimization algorithms) is also a speedup in iterations
 \begin{equation}\label{eq:26}
@@
 \end{equation}
-where $n(1)$ is the number of the trials carried out using the sequential method, and $n(p)$ is the number of the trials carried out using the parallel method with p processors.
-This characteristic is especially important since for solution of applied problems the time of carrying out of a trial exceeds the time of processing of its results.
+where $n(1)$ is the number of trials carried out using the sequential method, and $n(p)$ is the number of trials carried out using the parallel method with $p$ processors.
+This characteristic is especially important since for solving applied problems the time required to carry out a trial exceeds the time required to process its results.
 
-It is obvious that number of trials $n(p)$ for sequential and parallel algorithms described in sections \ref{sec:2} and \ref{sec:3} will differ.
+It is obvious that the number of trials $n(p)$ for sequential and parallel algorithms described in sections \ref{sec:2} and \ref{sec:3} will differ.
 Actually, the sequential algorithm of global search when selecting point $x^{k+1}$ of the next $k+1$ trial possesses complete information received at the previous $k$ iterations.
-The parallel algorithm of global search selects not one, but $p$ points $x^{k+j}, 1\leq j \leq p$,  at iteration $k+1$ based on the same information.
-It means that selection of point $x^{k+j}$ is carried out in absence of information on the trial results in points $x^{k+i}, 1\leq i<j$.
+Based on the same information, the parallel algorithm of global search selects $p$ points $x^{k+j}, 1\leq j \leq p$, rather than one at iteration $k+1$.
+It means that selection of point $x^{k+j}$ is carried out in absence of information on the trial results at points $x^{k+i}, 1\leq i<j$.
 Only the first point $x^{k+1}$ will coincide with the point selected by the sequential algorithm.
-Points of the other trials generally can not coincide with the points generated by the sequential algorithm.
-Carrying out of such trials can reduce efficiency of use of parallel processors.
+Other trials points generally cannot coincide with the points generated by the sequential algorithm.
+The efficiency offered by the use of parallel processors can be reduced when such trials are performed.
 Therefore, let us consider such trials as ``redundant'', and the value
@@
 \]
-as ``method redundancy''.
+will represent ``method redundancy''.
 
-Let us set the series of trials $\left\{x^k\right\}$ and $\left\{y^m\right\}$ generated correspondingly by sequential and parallel algorithms for solution of the same problem with $\epsilon = 0$ in the condition of stopping.
+Let us set a series of trials $\left\{x^k\right\}$ and $\left\{y^m\right\}$ generated correspondingly by sequential and parallel algorithms for solving the same problem with $\epsilon = 0$ in the stopping condition.
 The following theorem from \cite{RefStrongin2000} determines the number of computing elements $p$, which can be involved for non-redundant parallelization.
@@
 	\item
-	For value $M$ from (\ref{eq:11}) the following inequality
+	For the value $M$ from (\ref{eq:11}) the following inequality
 	\[
@@
 	\]
-	holds, where $H$ is H{\"o}lder constant of the minimized function.
+	holds, where $H$ is the H{\"o}lder constant of the minimized function.
 \end{enumerate}
@@
 
-The theorem corollary plays a special role for solution of multidimensional problems reduced to one-dimensional problems by means of Peano-like evolvent $y(x)$.
-Evolvent $y(x)$, which is approximation to Peano curve, has the effect of ``splitting'' of a point of the global minimum $y^\ast\in D$ to several preimages in interval $[0,1]$.
+The theorem corollary plays a special role in solving multidimensional problems reduced to one-dimensional problems by means of the Peano-like evolvent $y(x)$.
+The evolvent $y(x)$, which is an approximation to the Peano curve, has the effect of ``splitting'' of a global minimum point $y^\ast\in D$ to several preimages in the interval $[0,1]$.
 If function $\varphi(y)$ has the only global minimum in $D$, the ``reduced'' function $f(x)$ can have up to $2^N$ local extremum points close (by value) to a global extremum point (see \cite{RefStrongin2000}).
-In case of applying a parallel global search algorithm for minimization of a similar function it is possible receive non-redundancy when using up to $2^N+1$ computing elements.
+By applying a parallel global search algorithm for minimization of a similar function it is possible to achive non-redundancy when using up to $2^N+1$ computing elements.
 
-The parallel algorithm of global search showed a good speedup in implementation on CPU in systems with a small number of processors.
+The parallel algorithm of global search has demonstrated a good speedup in implementation on CPU in systems with a small number of processors.
 In \cite{RefGrishagin1997,RefStrongin2003} a speedup of algorithm close to linear was experimentally verified when carrying out 2-4 parallel trials.
-The results of work of the parallel GSA for minimization of 100 two-dimensional multiextremal functions described by formula (\ref{eq:19}) are shown in table \ref{tab:1}.
-All experiments used method parameters $r=2.9$, and $\epsilon=10^{-3}$, and the global minimum was found with the demanded accuracy in all $100$ problems.
+The results obtained with the parallel GSA for minimization of 100 two-dimensional multiextremal functions described by formula (\ref{eq:19}) are shown in table \ref{tab:1}.
+All experiments used method parameters $r=2.9$, and $\epsilon=10^{-3}$, and the global minimum was found with the required accuracy in all $100$ problems.
 The table shows the number of used processors $p$, the average number of trials $n(p)$ rounded to integers and average speedup in iterations $s(p)$ from (\ref{eq:26}).
@@
 
-What advantages does GPU have in case of implementation of global optimization algorithms?
+What are the advantages of GPU over CPU when implementing global optimization algorithms?
 First, CPU has only a small (up to $16$ in the advanced models) number of cores working independently at a high clock frequency.
@@
 GPU, on the contrary, works at low clock frequencies and uses hundreds of simpler computing elements.
-Second, a considerable share of a CPU crystal is occupied by a high-speed cache memory, while practically the whole of GPU consists of arithmetic and logic units.
+Second, a considerable part of a CPU crystal is occupied by a high-speed cache memory, while practically the whole of GPU consists of arithmetic and logic units.
 Therefore, GPUs are especially effective in problems, in which the number of arithmetic operations is great in comparison with memory operations.
 
-With respect to methods of global optimization, the operation which can be implemented effectively on GPU is simultaneous parallel computing of many values of the objective function.
-Naturally, for this purpose the function value calculation procedure has to be implemented on GPU.
-Transfers of information from CPU to GPU will be minimal: it is necessary only to transfer to GPU coordinates of trial points and to receive values of the function in these points.
-The functions that determine the processing of trial results according to algorithm and require work with a large volume of collected search information can be effectively implemented on CPU.
+With respect to methods of global optimization, parallel computing of many values of the objective function simultaneously is an operation which can be implemented effectively on GPU.
+In the function value calculation procedure implemented on GPU transfers of information from CPU to GPU will be minimal: it is necessary only to transfer the coordinates of trial points to GPU and to receive the values of the function in these points.
+The functions that determine the processing of trial results according to algorithm and require working with a large volume of collected search information can be effectively implemented on CPU.
 
-A general scheme of organization of calculations using GPU is shown in Fig.~\ref{fig:3}.
-According to this scheme steps 1 -- 4 of a parallel global search algorithm are carried out on a CPU.
+A general scheme showing the organization of calculations using GPU is given in Fig.~\ref{fig:3}.
+According to this scheme, steps 1 -- 4 of a parallel global search algorithm are carried out on a CPU.
 The coordinates of $p$ points of trials calculated at step 4 of the algorithm are collected in an intermediate buffer, and then are transferred to the graphics processing unit.
-Calculation of function values in these points is carried out on GPU, and then the results of trials (again via the intermediate buffer) are transferred to CPU.
+Calculation of function values at these points is carried out on a GPU, and then the results of trials (again via the intermediate buffer) are transferred to the CPU.
 
@@
 The use of several hundred of parallel computing elements (as in modern GPU) might not give a speedup of the search process by hundreds of times.
-In this case, conditions of the theorem of non-redundant parallelization can be violated: the number of local extremums will be less than the number of computing cores.
+In this case, the conditions of the theorem of non-redundant parallelization can be violated: the number of local extremums will be less than the number of computing cores.
 Then (according to the theorem and its corollary) a parallel GSA will generate redundant trial points.
-Nevertheless, despite certain redundancy, the general operating time of a GPU algorithm will be less than the operating time of a CPU algorithm.
+Nevertheless, despite certain redundancy, the overall operating time of a GPU algorithm will be less than the operating time of a CPU algorithm.
 This has been verified by computing experiments, the results of which are given in section \ref{sec:6}.
@@
 
-Keep in mind that widely known test problems from the domain of multidimensional global optimization are characterized by small time of objective function values calculation.
+It should be noted that widely known test problems from the field of multidimensional global optimization are characterized by small time required to calculate the objective function value.
 Usually, such a calculation is nothing else than a summation of several (according to problem dimension) values of elementary functions.
-For example, the well known Rastrigin’s test function, the global minimum of which is located in $y^\ast = (0,\dots,0)$ with $\varphi(y^\ast)=0$, is set by the expression
+For example, the well known Rastrigin test function, the global minimum of which is located in $y^\ast = (0,\dots,0)$ with $\varphi(y^\ast)=0$, is set by the expression
 \begin{equation}\label{eq:33}
@@
 
-Therefore, for the purpose of imitation of the computational complexity inherent to applied problems of optimization \cite{RefBarkalov2013}, calculation of the objective function in all performed experiments was complicated by additional calculations without changing the type of function and arrangement of its minimums (series summation from 20 thousand elements).
+Therefore, for the purpose of imitation of the computational complexity inherent to applied problems of optimization \cite{RefBarkalov2013}, calculation of the objective function in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (series summation from 20 thousand elements).
 
@@
 Threads in a thread block can interact effectively using information in shared memory and synchronization.
-For the purpose of assessing the importance of the size of a thread block during the period of execution of a GPU algorithm let us perform a preliminary series of experiments with the test function (\ref{eq:33}) with $N=5$.
+For the purpose of assessing the importance of the size of a thread block during the execution period of a GPU algorithm let us perform a preliminary series of experiments with the test function (\ref{eq:33}) with $N=5$.
 Let us vary the number $p$ of the used GPU threads and the size $pB$ of a thread block.
@@
 As is seen from this table, there is a speedup of the algorithm in the case of an increase in the number of trials at one iteration up to the number of cores on GPU, with subsequent saturation.
-Owing to a weak information dependence the size of the thread block $pB$ does not have determining influence, therefore let us use $pB=64$ in all subsequent experiments.
+Owing to a weak information dependence the size of the thread block $pB$ does not have decisive effect, therefore let us use $pB=64$ in all subsequent experiments.
 When carrying out this set of experiments the load on GPU was 15 -- 85 \% (depending on the number of threads); the load was measured by GPUZ program.
 
-After fixing the GPU parameters let us make the main computing experiments on a series of multidimensional multiextremal problems.
+After the parameters of the GPU algorithm have been determined let us make the main computing experiments on a series of multidimensional multiextremal problems.
 In work \cite{RefGaviano} a GKLS generator allowing generation of multiextremal optimization problems with known properties (number of local minima, size of their domains of attraction, global minimum point, function value, etc.) is described.
 
-The results of numerical comparison of the three sequential algorithms -- DIRECT \cite{RefJones}, DIRECT\textit{l} \cite{RefGablonsky} and global search algorithm (GSA) from section \ref{sec:2} -- are provided below (results of work of the first two algorithms are given in paper \cite{RefGaviano}).
+The results of numerical comparison of the three sequential algorithms -- DIRECT \cite{RefJones}, DIRECT\textit{l} \cite{RefGablonsky} and global search algorithm (GSA) from section \ref{sec:2} -- are provided below (results of the work of the first two algorithms are given in paper \cite{RefGaviano}).
 Numerical comparison was carried out on function classes \textit{Simple} and \textit{Hard} of dimension 4 and 5 from \cite{RefGaviano} since the solution of problems of dimension 2 and 3 demands a small number of iterations and it is impractical to use GPU for solving these problems.
@@
 
-Table \ref{tab:3} shows that GSA outperforms DIRECT and DIRECT\textit{l} methods on all classes of problems by average number of iterations.
-And in class 5-\textit{Hard} each of the methods solved not all problems: DIRECT did not solve 16 problems, DIRECT\textit{l} and GSA -- 4 problems.
+Table \ref{tab:3} shows that GSA outperforms DIRECT and DIRECT\textit{l} methods on all classes of problems in terms of average number of iterations.
+In class 5-\textit{Hard} all the three methods failed to solve all problems: DIRECT did not solve 16 problems, DIRECT\textit{l} and GSA, 4 problems.
 
 Let us estimate now the speedup when using the parallel global search algorithm implemented on CPU depending on the number $p$ of the used cores.
-Table \ref{tab:4} shows the average number of iterations performed by the method when solving a series of problems, and table \ref{tab:5} shows the average time requared to solve one problem.
-Tables \ref{tab:6} and \ref{tab:7} show speedup in time $S(p)$ and iterations $s(p)$ respectively; speedup of the parallel algorithm was measured in relation to the sequential one ($p=1$).
+Table \ref{tab:4} shows the average number of iterations performed by the method when solving a series of problems, and table \ref{tab:5} shows the average time required to solve one problem.
+Tables \ref{tab:6} and \ref{tab:7} show speedup in time $S(p)$ and iterations $s(p)$, respectively; speedup of the parallel algorithm was measured in relation to the sequential one ($p=1$).
 
@@
 
-The data show quintuple time speedup, when using 32 cores for all classes (except 5-\textit{Hard}).
-Delay in class 5-\textit{Hard} in transition from 16 to 32 cores is explained by an increase in average number of iterations and by the fact that out of 32 cores 16 are physical, and other 16 are virtual in hyper-threading mode, i.e.
-a real gain in performance in transition from 16 to 32 cores is less than double.
+The data show five-fold time speedup, when using 32 cores for all classes (except 5-\textit{Hard}).
+The lack of expected speedup in class 5-\textit{Hard} in transition from 16 to 32 cores is explained by an increase in the average number of iterations and by the fact that out of 32 cores 16 are physical, and other 16 are virtual in hyper-threading mode, i.e.
+a real gain in performance in transition from 16 to 32 cores is less than twofold.
 
 Let us make a similar experiment using a GPU version of the algorithm.
-Also let us fix the average number of iterations and average time spent on one problem (see the tables \ref{tab:last4} -- \ref{tab:last}).
+Also let us fix the average number of iterations and average time spent on one problem (see tables \ref{tab:last4} -- \ref{tab:last}).
 Let us measure the speedup and redundancy of GPU algorithm in relation to the CPU algorithm started with 32 cores (see column $p=32$ in tables \ref{tab:4} and \ref{tab:5}).
 The number of the used threads $p$ on GPU varied, all other parameters of the method did not change.
-The number of threads, which is greater, than the number of cores, was chosen to show the following effect: the number of iterations executed by an algorithm will be reduced in case of an increase in number of threads, while time of operation will not.
+The number of threads, which is greater, than the number of cores, was chosen to show the following effect: the number of iterations executed by the algorithm will be reduced in case of an increase in the number of threads, while the algorithm operation time will not be reduced.
 The minimum time of operation is reached, if the number of used threads does not surpass the number of GPU cores.
@@
 
-The results of experiments show advantage of the GPU algorithm in comparison with the CPU algorithm.
+The results of our experiments show some obvious advantages of the GPU algorithm in comparison with the CPU algorithm.
 The algorithm using the maximum quantity of cores of CPU works slower than any of the given GPU implementations.
 In addition, the results show a considerable speedup in the number of iterations.
-For example, solution of problems of the most difficult of the classes considered -- 5-\textit{Hard} -- required on average only 353 parallel iterations on GPU, whereas even when using all computing CPU cores the number of iterations made was over eleven thousand.
+For example, on average only 353 parallel iterations on GPU were sufficient to solve problems of the most difficult class considered -- 5-\textit{Hard}, whereas even with the use of all computing CPU cores over eleven thousand iterations were required.
 Finally, the GPU algorithm is almost non-redundant in relation to the CPU algorithm.
@@
 
-By now, parallel implementations of the majority of the developed sequential algorithms have been suggested.
-Please find a review of modern methods of parallel global optimization, for example, in \cite{RefApuzzo}.
+By now, parallel implementations have been suggested for the most of the existing sequential algorithms.
+A review of modern methods of parallel global optimization can be found, for example, in \cite{RefApuzzo}.
 
-The methods described in \cite{RefApuzzo} can be divided into two classes: deterministic methods of global optimization (including the algorithm of global search considered in this article), and heuristic methods, which are one way or another based on the concept random search.
-Algorithms belonging to the latter class provide for use of massive parallelism and are easily implemented on GPU (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi} and the review \cite{RefLangdon}).
+The methods described in \cite{RefApuzzo} can be divided into two classes: deterministic methods of global optimization (including the algorithm of global search considered in this article), and heuristic methods, which are one way or another based on the random search concept.
+Algorithms belonging to the latter class allow use of massive parallelism and are easily implemented on GPU (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi} and the review \cite{RefLangdon}).
 
@@
 For example, parallelization of algorithms based on the ideas of the branch and bound method is carried out for a small number of processors using OpenMP and MPI \cite{RefPaulavicius}.
-It is noteworthy that the branch and bound method is applied in other sections of optimization; in particular, combinatorial optimization and integer programming.
-Using GPU shows good results in these domains (see \cite{RefBoukedjar}, \cite{RefCarneiro}).
+It is noteworthy that the branch and bound method is applied in other areas of optimization; in particular, combinatorial optimization and integer programming.
+Using GPU shows good results in these areas (see \cite{RefBoukedjar}, \cite{RefCarneiro}).
 It is explained by the following factor.
-Problems of this class (such as travelling salesman problem or knapsack problem) are characterized by short time of criterion value calculation in one point and, at the same time, relatively long time of results processing.
+Problems of this class (such as the Travelling Salesman Problem or the Knapsack problem) are characterized by short time of criterion value calculation in one point and, at the same time, relatively long time required to process the results.
 Therefore, the most labor-intensive operations here are branching and bound computation, which are implemented on GPU (see \cite{RefBoukedjar}, \cite{RefCarneiro}).
 
-This paper considers the problems of lipshitzian global optimization, that are frequently met in applications.
-They are characterized by a considerable time of calculation of function value in one point (in comparison with processing of results of such calculation).
-Therefore, calculation of the value of an optimized function in one point is done via GPU, and the main objective of the algorithm (working at CPU) is to efficiently select points of parallel trials.
+This paper considers the problems of lipschitzian global optimization that are frequently encountered in applications.
+They are characterized by a considerable time required to calculate the function value in one point (in comparison with the processing of the results of such calculation).
+Therefore, calculation of the value of an optimized function in one point is done using a GPU, and the main objective of the algorithm (working at CPU) is to efficiently select points of parallel trials.
 This is carried out in the parallel implementation of GSA on GPU.
 
-Summing up the results, let us note that use of graphics processing units for solution of global optimization problems is a new, promising direction of research as high efficiency of modern supercomputers is reached (generally) due to use of graphics accelerators.
+Summing up the results, let us note that use of graphics processing units for solving global optimization problems is a new, promising line of research as high efficiency of modern supercomputers is reached (generally) due to the use of graphics accelerators.
 
-This paper considers a parallel algorithm of solution of multidimensional multiextremal optimization problems and its implementation on GPU.
+This paper considers a parallel algorithm for solving multidimensional multiextremal optimization problems and its implementation on GPU.
 This algorithm was developed within the framework of information-statistical approach.
-Theoretical information on properties of parallel algorithm characterizing conditions of non-redundant parallelization are provided.
+Theoretical information on the properties of the parallel algorithm characterizing the conditions of non-redundant parallelization are provided.
 
-For the purpose of experimental demonstration of theoretical properties of the considered parallel algorithms, computing experiments on a series of several hundred test problems of different dimension were performed.
-The results of work of the sequential global search algorithm show its superiority over other known methods of global optimization.
+For the purpose of experimental demonstration of theoretical properties of the considered parallel algorithms, computing experiments on a series of several hundred test problems of different dimensions were performed.
+The results obtained with the sequential global search algorithm show its superiority over other known methods of global optimization.
 At the same time the parallel algorithm shows a good speedup and low redundancy (both GPU and CPU version), and the GPU version of the parallel algorithm shows a speedup of 2 to 6 times in comparison with the fastest parallel CPU version.

### 2019-10-07 2016/JoGO_GPU/BarkalovGergel.tex
JOGO2016 paper next version

@@
 
-This present paper considers parallel algorithms for solving multiextremal optimization problems.
+This paper considers parallel algorithms for solving multiextremal optimization problems.
 The objective function in such problems is characterized by existence of several local optima (usually their number is not known and can be large).
-In multiextremal problems the possibility of a reliable estimation of the global optimum is essentially based on the availability of an a priori information about the function, allowing us to link possible values of an optimizable function with known values at search points.
+%new text begin
+In multiextremal problems the possibility of a reliable estimation of the global optimum is essentially based on the availability of an a priori information about the function, allowing us to estimate lower bound of the objective function using known values at search points.
+%new text end
 Very often, such information about the problem is presented as the assumption that the objective function $\varphi$ satisfies a Lipschitz condition with an a priori unknown constant $L$. 
-The objective function can be represented by the algorithm for calculating its values in the search domain, i.e.
-it can be a ``black-box'' function.
-Similar problems are common in applications (problems of optimal design of objects and technological processes in various fields of technology, problems of model fitting according to observed data in scientific research, optimization of control processes, etc., see the review \cite{RefPinter}).
+%new text begin
+The important feature of the problem under consideration is that the objective function is commonly not given by an analytical formula.
+The usual way to calculate the function value is to apply a computer-aided algorithm.
+Thus, each calculation may be interpreted as running some ``black-box'' to produce the outcome for the preset input.
+These runs could require substantial computer resources and therefore should not be too numerous.
+%new text end
+%new word - such
+Such problems are common in applications (problems of optimal design of objects and technological processes in various fields of technology, problems of model fitting according to observed data in scientific research, optimization of control processes, etc., see the review \cite{RefPinter}).
 
@@
 Furthermore, some principles of development of graphics hardware have changed over the past few years.
-Now graphics hardware is more programmable.
 Today a GPU is a high-performance flexibly programmable and massively parallel processor that can provide solution to many computationally complex problems \cite{RefHwu}.
 
-However, in the field of global optimization the potential of GPUs is not completely utilized.
-GPUs are generally used for parallelization of algorithms, which one way or another are based on the concept of random search (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi}); a review of this trend of GPU optimization algorithms is given in \cite{RefLangdon}.
+%new text begin
+By now, parallel implementations have been suggested for most of the existing sequential algorithms.
+A review of modern methods of parallel global optimization can be found, for example, in \cite{RefApuzzo}.
+The methods described in \cite{RefApuzzo} can be divided into two classes: deterministic methods of global optimization, and heuristic methods, which are one way or another based on the random search concept.
+Algorithms belonging to the latter class allow the use of massive parallelism and are easily implemented on GPU (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi} and the review \cite{RefLangdon}).
 Owing to their stochastic nature, algorithms of this type generally do not guarantee convergence to a globally optimal solution, and that sets them apart from deterministic methods.
 
-Some parallel variants are proposed for many deterministic algorithms of lipschitzian global optimization with guaranteed convergence \cite{RefGergel2005,RefEvtushenko,RefHe,RefPaulavicius}.
-However, these versions of algorithms are parallelized on CPU using MPI and/or OpenMP technologies; there are no implementations on GPU.
-The present article contains the results of an investigation into the GPU implementation of the global search parallel algorithm developed according to the information-statistical approach presented in the monograph \cite{RefStrongin2000}.
+Some parallel variants are proposed for deterministic algorithms of lipschitzian global optimization with guaranteed convergence \cite{RefGergel2005,RefEvtushenko,RefHe,RefPaulavicius}.
+However, these versions of algorithms are parallelized on CPU; there are no implementations on GPU.
+For example, parallelization of algorithms based on the ideas of the branch and bound method is carried out for a small number of processors using OpenMP and MPI technologies \cite{RefPaulavicius}.
+It is noteworthy that the branch and bound method is applied in other areas of optimization; in particular, combinatorial optimization and integer programming.
+Using GPU shows good results in these areas \cite{RefBoukedjar}, \cite{RefCarneiro}.
+It is explained by the following factor.
+Problems of this class (such as the Travelling Salesman Problem or the Knapsack Problem) are characterized by short time of objective function value calculation in one point and, at the same time, relatively long time required to process the results.
+Therefore, the most labor-intensive operations here are branching and bound computation, which are implemented on GPU (see \cite{RefBoukedjar}, \cite{RefCarneiro}).
+
+This paper considers the problems of lipschitzian global optimization that are frequently encountered in applications.
+They are characterized by a considerable time required to calculate the function value in one point (in comparison with the processing of the results of such calculation).
+Therefore, calculation of the value of an optimized function in one point can be done using a GPU, and the main objective of the algorithm (working at CPU) is to efficiently select points of parallel trials.
+The article contains the results of an investigation into the GPU implementation of the global search parallel algorithm developed according to the information-statistical approach presented in the monograph \cite{RefStrongin2000}.
+%new text end
 
@@
 \]
+%new text begin
 Problems of numerical construction of Peano-type space filling curves and the corresponding theory are considered in detail in \cite{RefStrongin2000}, \cite{RefSergeyev2013}.  
-Here we will note that a numerically constructed curve (\textit{evolvent}) is an approximation to a theoretical Peano curve with an accuracy $2^{-m}$ , where $m$ is an evolvent construction parameter.
+%new text begin
+Here we will note that a numerically constructed curve (\textit{evolvent}) is $2^{-m}$ accurate approximation of the theoretical Peano curve in $L_\infty$ metric, where $m$ is an evolvent construction parameter.
+Examples of the evolvent with different $m$ in two dimensions are given in Fig.~\ref{fig:0}.
+
+%Here we will note that a numerically constructed curve (\textit{evolvent}) is an approximation to a theoretical Peano curve with an accuracy $2^{-m}$ , where $m$ is an evolvent construction parameter.
 It means that using an evolvent we can distinguish points in a multidimensional space, the coordinates of which differ by not less than $2^{-m}$; closer points are considered as coincident.
-Thereby an evolvent constructed with a fixed m allows obtaining a solution of a problem with an accuracy of not more than $2^{-m}$ by coordinate.
+Thereby an evolvent constructed with a fixed $m$ allows obtaining a solution of a problem with an accuracy of not more than $2^{-m}$ by coordinate.
+
+\begin{figure}
+\begin{minipage}{0.32\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1a.JPG} \\ (a)}
+\end{minipage}
+\hfill
+\begin{minipage}{0.32\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (b)}
+\end{minipage}
+\begin{minipage}{0.32\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (c)}
+\end{minipage}
+\caption{Elvolvents in two dimensions with (a) $m=3$, (b) $m=4$ and (c) $m=5$}
+\label{fig:0}
+\end{figure}
+%new text end
 
@@
 satisfying the H{\"o}lder condition.
+%new text begin
+Once again, we recall that the Peano curve $y(x)$ is defined as a limit object emerging in some sequential construction. Appropriate approximations to $y(x)$ are to be used for practical application.
+%new text end
 
@@
 
-The algorithm terminates if the condition $\Delta_t<\epsilon$ is satisfied; here $\epsilon>0$ the accuracy is preset. For estimation of the global solution, values
+The algorithm terminates if the condition $\Delta_t<\epsilon$ is satisfied; here $\epsilon>0$ is the preset accuracy. For estimation of the global solution, values
 \[
@@
 
+%new text begin
 Remark 1.
-The normalized characteristics $R(i)$ from (\ref{eq:14}) can be viewed as probabilities of locating a global minimum within the interval $[x_{i-1},x_i], 1\leq i \leq k+1$.
-This statement is proved in \cite{RefStrongin2000}.
-Thus, at each iteration a new trial point is selected inside the interval, which has the greatest probability of finding a global minimum.
+The global search algorithm was developed in the framework of a stochastic model (see \cite{RefStrongin2000}).
+From this point of view the normalized characteristics $R(i)$ from (\ref{eq:14}) can be considered as probabilities of locating the global minimum within the interval  $(x_{i-1},x_i), 1\leq i \leq k+1$.
+Thus, at each iteration a new trial point is selected inside the interval, which has the greatest probability of finding the global minimum.
+%new text end
 
@@
 	\center
-  \includegraphics[width=0.75\textwidth]{fig1.jpg} 
+  \includegraphics[width=0.75\textwidth]{fig2.jpg} 
   \caption{Minimization of a two-dimensional function by the sequential algorithm}
@@
 
-The first option is to parallelize calculation of the target function describing the optimized object. This method can provide some speedup, but is specific for each problem.
+The first option is to parallelize calculation of the objective function describing the optimized object. This method can provide some speedup, but is specific for each problem.
 
@@
 
-The third option is to  change the scheme of an algorithm for the purpose of carrying out several trials in parallel.
-This approach is most promising since it is characterized by efficiency (the part of computing process, in which the bulk of calculations is carried out, is parallelized) and generality (it is applicable for a wide class of characteristical algorithms of multiextremal optimization \cite{RefGrishagin1997}).
+%new text begin
+The third option is to change an algorithm for the purpose of carrying out several trials in parallel.
+This approach is most promising since it is characterized by efficiency and generality.
+Firstly, the part of computing process, in which the bulk of calculations is carried out, is parallelized.
+Secondly, the approach is applicable for a wide class of characteristical algorithms of multiextremal optimization \cite{RefGrishagin1997}.
+%new text end
 
@@
 
-Steps 1--3 of the parallel algorithm fully repeat steps 1--3 of the global search algorithm.
+Steps 1--3 of the parallel algorithm fully repeat steps 1--3 of the sequential global search algorithm.
 
@@
 	\center
-  \includegraphics[width=0.75\textwidth]{fig2.jpg} 
+  \includegraphics[width=0.75\textwidth]{fig3.jpg} 
   \caption{Minimization of a two-dimensional function by the parallel algorithm}
@@
 First, CPU has only a small (up to $16$ in the advanced models) number of cores working independently at a high clock frequency.
-Each CPU core is a powerful computer.
 GPU, on the contrary, works at low clock frequencies and uses hundreds of simpler computing elements.
@@
 
-With respect to methods of global optimization, parallel computing of many values of the objective function simultaneously is an operation which can be implemented effectively on GPU.
+With respect to methods of global optimization, parallel computing of many values of the objective function simultaneously is an operation which can be efficiently implemented on GPU.
 In the function value calculation procedure implemented on GPU transfers of information from CPU to GPU will be minimal: it is necessary only to transfer the coordinates of trial points to GPU and to receive the values of the function in these points.
-The functions that determine the processing of trial results according to algorithm and require working with a large volume of collected search information can be effectively implemented on CPU.
+The functions that determine the processing of trial results according to algorithm and require working with a large volume of collected search information can be efficiently implemented on CPU.
 
@@
 	\center
-  \includegraphics[width=0.75\textwidth]{fig3.jpg} 
+  \includegraphics[width=0.75\textwidth]{fig4.jpg} 
   \caption{Scheme of information exchanges in a GPU algorithm}
@@
 For the purpose of assessing the importance of the size of a thread block during the execution period of a GPU algorithm let us perform a preliminary series of experiments with the test function (\ref{eq:33}) with $N=5$.
-Let us vary the number $p$ of the used GPU threads and the size $pB$ of a thread block.
+Let us vary the number $p$ of the used GPU threads and the size $p_b$ of a thread block.
 All the other parameters of the method do not change ($r=2.3$, $\epsilon = 10^{-3}$, the evolvent construction parameter $m=10$).
-Table \ref{tab:2} shows the GPU algorithm operating time depending on $p$ and $pB$.
+Table \ref{tab:2} shows the GPU algorithm operating time depending on $p$ and $p_b$.
 \begin{table}
@@
 		\hline\noalign{\smallskip}
-%		 $p$ & $pB=16$ & $pB=32$ & $pB=64$ & $pB=128$ & $pB=256$  \\
-		$p$ & \multicolumn{5}{l}{ \hfil $pB$ \hfil } \\
+%		 $p$ & $p_b=16$ & $p_b=32$ & $p_b=64$ & $p_b=128$ & $p_b=256$  \\
+		$p$ & \multicolumn{5}{l}{ \hfil $p_b$ \hfil } \\
 		\noalign{\smallskip} \cline{2-6} \noalign{\smallskip}
@@
 As is seen from this table, there is a speedup of the algorithm in the case of an increase in the number of trials at one iteration up to the number of cores on GPU, with subsequent saturation.
-Owing to a weak information dependence the size of the thread block $pB$ does not have decisive effect, therefore let us use $pB=64$ in all subsequent experiments.
+Owing to a weak information dependence the size of the thread block $p_b$ does not have decisive effect, therefore let us use $p_b=64$ in all subsequent experiments.
 When carrying out this set of experiments the load on GPU was 15 -- 85 \% (depending on the number of threads); the load was measured by GPUZ program.
@@
 After the parameters of the GPU algorithm have been determined let us make the main computing experiments on a series of multidimensional multiextremal problems.
-In work \cite{RefGaviano} a GKLS generator allowing generation of multiextremal optimization problems with known properties (number of local minima, size of their domains of attraction, global minimum point, function value, etc.) is described.
+In work \cite{RefGaviano} a GKLS generator allowing generation of multiextremal optimization problems with known properties (number of local minima, size of their domains of attraction, global minimum point, etc.) is described.
 
 The results of numerical comparison of the three sequential algorithms -- DIRECT \cite{RefJones}, DIRECT\textit{l} \cite{RefGablonsky} and global search algorithm (GSA) from section \ref{sec:2} -- are provided below (results of the work of the first two algorithms are given in paper \cite{RefGaviano}).
-Numerical comparison was carried out on function classes \textit{Simple} and \textit{Hard} of dimension 4 and 5 from \cite{RefGaviano} since the solution of problems of dimension 2 and 3 demands a small number of iterations and it is impractical to use GPU for solving these problems.
+Numerical comparison was carried out on function classes \textit{Simple} and \textit{Hard} of dimension 4 and 5 from \cite{RefGaviano} since solving a problems of dimension 2 and 3 demands a small number of iterations and it is impractical to use GPU for solving these problems.
 Global minimum $y^\ast$ was considered to be found, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e.
@@
 The size of the vicinity was selected (according to \cite{RefGaviano}) as $\delta = \left\|b-a\right\|\sqrt[N]{\Delta}$, $N$ -- problem dimension, $a$ and $b$ -- borders of search domain $D$, parameter $\Delta=10^{-6}$ at $N=4$ and $\Delta=10^{-7}$ at $N=5$.
-When using the GSA for the class \textit{Simple} parameter $r=4.5$ was selected, for the class \textit{Hard} $r=5.6$ was selected; Peano curve construction parameter was fixed as $m=10$.
+When using the GSA for the class \textit{Simple} parameter $r=4.5$ was selected, for the class \textit{Hard} $r=5.6$ was selected; evolvent construction parameter was fixed as $m=10$.
 The maximum allowable number of iterations was $K_{max} = 10^6$.
@@
 The average number of iterations $k_{av}$ performed by the method for solving a series of problems from these classes is shown in table \ref{tab:3}.
-Symbol $>$ reflects a situation in which not all problems of the class were solved by the method.
+Symbol ''$>$''  reflects a situation in which not all problems of the class were solved by the method.
 This means that the algorithm was stopped as the maximum allowable number of iterations $K_{max}$ was achieved.
@@
 			5	& \textit{Simple} &	$>$16057(1) &	16758 &	15920 \\
-				& \textit{Hard} &	$>$217215(16) &	$>$269064(4) & $>$148342 (4) \\
+				& \textit{Hard} &	$>$217215(16) &	$>$269064(4) & $>$148342(4) \\
 		\noalign{\smallskip}\hline
@@
 Let us make a similar experiment using a GPU version of the algorithm.
-Also let us fix the average number of iterations and average time spent on one problem (see tables \ref{tab:last4} -- \ref{tab:last}).
+Also let us fix the average number of iterations and average time required to solve one problem (see tables \ref{tab:last4} -- \ref{tab:last}).
 Let us measure the speedup and redundancy of GPU algorithm in relation to the CPU algorithm started with 32 cores (see column $p=32$ in tables \ref{tab:4} and \ref{tab:5}).
 The number of the used threads $p$ on GPU varied, all other parameters of the method did not change.
-The number of threads, which is greater, than the number of cores, was chosen to show the following effect: the number of iterations executed by the algorithm will be reduced in case of an increase in the number of threads, while the algorithm operation time will not be reduced.
+The number of threads, which is greater, than the number of cores, was chosen to show the following effect: the number of iterations executed by the algorithm will be reduced in case of an increase of the threads number, while the algorithm operation time will not be reduced.
 The minimum time of operation is reached, if the number of used threads does not surpass the number of GPU cores.
@@
 
-By now, parallel implementations have been suggested for the most of the existing sequential algorithms.
-A review of modern methods of parallel global optimization can be found, for example, in \cite{RefApuzzo}.
-
-The methods described in \cite{RefApuzzo} can be divided into two classes: deterministic methods of global optimization (including the algorithm of global search considered in this article), and heuristic methods, which are one way or another based on the random search concept.
-Algorithms belonging to the latter class allow use of massive parallelism and are easily implemented on GPU (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi} and the review \cite{RefLangdon}).
-
-Global optimization of implementations on GPU for deterministic algorithms has not been suggested yet.
-For example, parallelization of algorithms based on the ideas of the branch and bound method is carried out for a small number of processors using OpenMP and MPI \cite{RefPaulavicius}.
-It is noteworthy that the branch and bound method is applied in other areas of optimization; in particular, combinatorial optimization and integer programming.
-Using GPU shows good results in these areas (see \cite{RefBoukedjar}, \cite{RefCarneiro}).
-It is explained by the following factor.
-Problems of this class (such as the Travelling Salesman Problem or the Knapsack problem) are characterized by short time of criterion value calculation in one point and, at the same time, relatively long time required to process the results.
-Therefore, the most labor-intensive operations here are branching and bound computation, which are implemented on GPU (see \cite{RefBoukedjar}, \cite{RefCarneiro}).
-
-This paper considers the problems of lipschitzian global optimization that are frequently encountered in applications.
-They are characterized by a considerable time required to calculate the function value in one point (in comparison with the processing of the results of such calculation).
-Therefore, calculation of the value of an optimized function in one point is done using a GPU, and the main objective of the algorithm (working at CPU) is to efficiently select points of parallel trials.
-This is carried out in the parallel implementation of GSA on GPU.
-
 Summing up the results, let us note that use of graphics processing units for solving global optimization problems is a new, promising line of research as high efficiency of modern supercomputers is reached (generally) due to the use of graphics accelerators.
@@
 
+\bibitem{RefApuzzo}
+D'Apuzzo, M.,  Marino, M., Migdalas, A., Pardalos, P.M., Toraldo, G.: Parallel computing in global optimization.
+In Handbook of Parallel Computing and Statistics, Chapman \& Hall, London, 225--258 (2006)
+
 \bibitem{RefFerreiro}
@@
 Paulavicius, R., Zilinskas, J., Grothey, A.: Parallel branch and bound for global optimization with combination of Lipschitz bounds. Optimization Methods \& Software 26(3), 487--498, (2011)
+
+\bibitem{RefBoukedjar}
+Boukedjar A., Lalami M. E., El Baz D.: Parallel Branch and Bound on a CPU-GPU System. In 20th Int. Conf. on Parallel,  Distributed and Network-Based Processing, 392--398 (2012)
+
+\bibitem{RefCarneiro}
+Carneiro, T. et al.: A new parallel schema for branch-and-bound algorithms using GPGPU. In 23rd Int. Symposium on Computer Architecture and High Performance Computing, 41--47 (2011)
 
@@
 
-\bibitem{RefApuzzo}
-M.
-D’Apuzzo, M.
-Marino, A.
-Migdalas, P.M.
-Pardalos, and G.
-Toraldo, Parallel computing in global optimization, in Handbook of Parallel Computing and Statistics, E.J.
-Kontoghiorghes, ed.
-Chapman \& Hall, London, 2006, pp.
-225–258
-
-\bibitem{RefBoukedjar}
-Boukedjar A., Lalami M. E., El Baz D. Parallel Branch and Bound on a CPU-GPU System //PDP. – 2012. – С. 392-398
-
-\bibitem{RefCarneiro}
-Carneiro T.
-et al.
-A new parallel schema for branch-and-bound algorithms using GPGPU //Computer Architecture and High Performance Computing (SBAC-PAD), 2011 23rd International Symposium on.
-– IEEE, 2011.
-– С.
-41-47.
-
 \end{thebibliography}

### 2019-10-07 2016/JoGO_GPU/BarkalovGergel.tex
JOGO2016 paper next version

@@
 \title{Parallel global optimization on GPU 
-	\thanks{The research is supported by the grant of the Ministry of education and science of the Russian Federation (the agreement of August 27, 2013 No 02.B.49.21.0003)}
+	\thanks{The revised version of the paper was supported by the Russian Science Foundation, project No 15-11-30022 ``Global optimization, supercomputing computations, and applications''}
 }
@@
 \institute{K. Barkalov  \at
-              Computing Mathematics and Cybernetics Faculty, Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
+              Computational Mathematics and Cybernetics Faculty, Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
               Tel.: +7-831-4623356\\
@@
 Conditions of non-redundant parallelization of the algorithm are considered.
-Implementation of the algorithm on modern computing systems with the use of a graphics processing unit (GPU) is investigated.
+Efficiency of the algorithm on modern computing systems with the use of graphics processing units (GPUs) is investigated.
 Speedup of the algorithm using GPU as compared with the same algorithm implemented on CPU only is demonstrated experimentally.
-Computing experiments are carried out on a series of several hundred multidimensional multiextremal problems.
+Computational experiments are carried out on a series of several hundred multidimensional multiextremal problems.
 \keywords{Global optimization \and Parallel computing \and Graphics processing unit \and Peano-type space filling curves \and Characteristical algorithms}
@@
 This paper considers parallel algorithms for solving multiextremal optimization problems.
-The objective function in such problems is characterized by existence of several local optima (usually their number is not known and can be large).
+The objective function in such problems has several local optima (usually their number is not known and can be large).
 %new text begin
-In multiextremal problems the possibility of a reliable estimation of the global optimum is essentially based on the availability of an a priori information about the function, allowing us to estimate lower bound of the objective function using known values at search points.
+In such problems a reliable estimation of the global optimum is possible only with an a priori information about the function, allowing to estimate lower bound of the objective function using known values at search points.
 %new text end
-Very often, such information about the problem is presented as the assumption that the objective function $\varphi$ satisfies a Lipschitz condition with an a priori unknown constant $L$. 
+Very often, such information is presented as the assumption that the objective function satisfies a Lipschitz condition with an a priori unknown constant $L$. 
 %new text begin
-The important feature of the problem under consideration is that the objective function is commonly not given by an analytical formula.
+The important feature of the global optimization problems is that the objective function is commonly not given by an analytical formula.
 The usual way to calculate the function value is to apply a computer-aided algorithm.
-Thus, each calculation may be interpreted as running some ``black-box'' to produce the outcome for the preset input.
+Thus, each such calculation may be interpreted as running some blackbox device  to produce the outcome for the preset input.
 These runs could require substantial computer resources and therefore should not be too numerous.
@@
 Multiextremal optimization problems offer little in terms of analytical solution, while numerical solution is computationally challenging as computational costs often grow exponentially in dimension.
-By employing modern parallel computational systems it is possible to broaden the area of applicability of global optimization methods.
+Using modern parallel computational systems it is possible to broaden the area of applicability of global optimization methods.
 At the same time, the problem of efficient parallelization of the global search process arises.
-Therefore, development of efficient parallel methods for numerical solution of multiextremal optimization problems and their software implementation for modern computational systems is the obvious focus of research in this area.
+Therefore, development of efficient parallel methods for numerical solution of multiextremal optimization problems and implementation of corresponding software for modern computational systems is the obvious focus of research in this area.
 
 One of the promising trends in the field of parallel global optimization (as in many areas connected with software implementation of complex algorithms) is the use of graphics processing units (GPUs).
-During the last decade, productivity of GPUs has been rapidly growing to satisfy the continuously growing demand of graphic application developers.
+During the last decade, productivity of GPUs has been rapidly increasing to satisfy the continuously growing demand of graphic application developers.
 Furthermore, some principles of development of graphics hardware have changed over the past few years.
-Today a GPU is a high-performance flexibly programmable and massively parallel processor that can provide solution to many computationally complex problems \cite{RefHwu}.
+Today a GPU is a high-performance flexibly programmable and massively parallel processor that can provide solution to many computationally intensive problems \cite{RefHwu}.
 
@@
 A review of modern methods of parallel global optimization can be found, for example, in \cite{RefApuzzo}.
-The methods described in \cite{RefApuzzo} can be divided into two classes: deterministic methods of global optimization, and heuristic methods, which are one way or another based on the random search concept.
-Algorithms belonging to the latter class allow the use of massive parallelism and are easily implemented on GPU (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi} and the review \cite{RefLangdon}).
-Owing to their stochastic nature, algorithms of this type generally do not guarantee convergence to a globally optimal solution, and that sets them apart from deterministic methods.
+The methods described in \cite{RefApuzzo} can be divided into two classes: deterministic methods of global optimization, and heuristic methods, which are based one way or another on the random search concept.
+Algorithms, which belong to the latter class, allow the use of massive parallelism and are easily implemented on GPU (see works \cite{RefFerreiro,RefZhu,RefGarcia,RefMussi} and the review \cite{RefLangdon}).
+Owing to their stochastic nature, algorithms of this type can guarantee convergence to a globally optimal solution only in probabilistic sense, and that sets them apart from deterministic methods.
 
-Some parallel variants are proposed for deterministic algorithms of lipschitzian global optimization with guaranteed convergence \cite{RefGergel2005,RefEvtushenko,RefHe,RefPaulavicius}.
-However, these versions of algorithms are parallelized on CPU; there are no implementations on GPU.
-For example, parallelization of algorithms based on the ideas of the branch and bound method is carried out for a small number of processors using OpenMP and MPI technologies \cite{RefPaulavicius}.
+Some parallel variants are proposed for deterministic algorithms of Lipschitzian global optimization in \cite{RefGergel2005,RefEvtushenko,RefHe,RefPaulavicius}.
+However, these versions of algorithms are parallelized on a computing cluster with distributed memory using multi-core CPU; implementations of Lipschitzian global optimization algorithms on GPU have not been suggested yet.
+For example, parallelization of algorithms based on the ideas of the branch and bound method is carried out for a small number of nodes using OpenMP and MPI technologies \cite{RefPaulavicius}.
 It is noteworthy that the branch and bound method is applied in other areas of optimization; in particular, combinatorial optimization and integer programming.
-Using GPU shows good results in these areas \cite{RefBoukedjar}, \cite{RefCarneiro}.
+Using GPU shows good results in these areas \cite{RefBoukedjar,RefCarneiro}.
 It is explained by the following factor.
-Problems of this class (such as the Travelling Salesman Problem or the Knapsack Problem) are characterized by short time of objective function value calculation in one point and, at the same time, relatively long time required to process the results.
-Therefore, the most labor-intensive operations here are branching and bound computation, which are implemented on GPU (see \cite{RefBoukedjar}, \cite{RefCarneiro}).
+Problems of this class (such as the Travelling Salesman Problem or the Knapsack Problem) are characterized by short time of objective function value calculation at one point and, at the same time, relatively long time required to process the results.
+Therefore, the most of labor-intensive operations here consist in branching and bound computations, which are implemented on GPU (see \cite{RefBoukedjar,RefCarneiro}).
 
-This paper considers the problems of lipschitzian global optimization that are frequently encountered in applications.
-They are characterized by a considerable time required to calculate the function value in one point (in comparison with the processing of the results of such calculation).
-Therefore, calculation of the value of an optimized function in one point can be done using a GPU, and the main objective of the algorithm (working at CPU) is to efficiently select points of parallel trials.
+This paper considers the problems of Lipschitzian global optimization that are frequently encountered in applications.
+They are characterized by a considerable time required to calculate the function value at one point (in comparison with the processing of the results of such calculation).
+For example, the objective function can be defined by solving SLAE, ODE, PDE, or Monte Carlo simulation.
+Currently, GPU may be used to solve all these types of problems \cite{RefKindratenko}.
+Therefore, calculation of values of an optimized function can be done using a GPU, and the main objective of the algorithm (working at CPU) is to efficiently select points of parallel trials.
 The article contains the results of an investigation into the GPU implementation of the global search parallel algorithm developed according to the information-statistical approach presented in the monograph \cite{RefStrongin2000}.
+Some preliminary results of research in this area have been published in \cite{RefLebedevBarkalov}.
 %new text end
@@
 Section \ref{sec:2} contains a description of the basic algorithm of global search.
-Section \ref{sec:3} contains a scheme of its parallelization, and section \ref{sec:4} includes theoretical statements that characterize the speedup and non-redundancy of the parallel algorithm.
+Section \ref{sec:3} describes a scheme of its parallelization, and section \ref{sec:4} includes theoretical statements that characterize the speedup and non-redundancy of the parallel algorithm.
 In section \ref{sec:5}, ways of possible implementation of the parallel algorithm on GPU are discussed.
@@
 
-Let us consider the problem of search for a global minimum of an $N$-dimensional function $\varphi(y)$ within a hyper interval $D$
+Let us consider the problem of search for a global minimum of an $N$-dimensional function $\varphi(y)$ within a hyperinterval $D$
 \begin{eqnarray}\label{eq:1}
@@
 
-By using this kind of mapping it is possible to reduce the multidimensional problem~(\ref{eq:1}) to a one-dimensional problem
+By using this kind of mapping it is possible to reduce the multidimensional problem~(\ref{eq:1}) to a univariate problem
 \[
@@
 
-Therefore, it is possible, without limitation of generality, to consider minimization of one-dimensional function
+Therefore, it is possible, without loss of generality, to consider minimization of univariate function
 \[
@@
 At the first iteration of the method the trial is carried out at an arbitrary internal point $x^1$ of the interval $[0,1]$.
-The point of trial at the next iteration $(k+1)$ is determined according to these rules.
+The point of trial at the next iteration $(k+1)$ is determined according to the rules presented below.
 
@@
 \]
-which includes boundary points of the interval $[0,1]$ and the points of the previous trials, with subscripts in increasing order of coordinate values, i.e.
+which includes boundary points of the interval $[0,1]$ and the points of the previous trials, with subscripts in increasing order of coordinate values, i.e.,
 \[
@@
 
-Step 3. Calculate a \textit{characteristic} for every interval $(x_{i-1}, x_i), \; 1\leq i \leq k+1,$   according to the following formulas
+Step 3. Calculate a \textit{characteristic} for every interval $(x_{i-1}, x_i), \; 1\leq i \leq k+1,$   according to the following formulae
 \[
@@
 Step 4. Find interval $(x_{t-1},x_t)$ with the maximum characteristic
-\[
+\begin{equation}\label{eq:141}
 R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
-\]
+\end{equation}
 
-Step 5. Carry out a trial at the point $x^{k+1}\in(x_{t-1},x_t)$, calculated using the following formulas
+Step 5. Carry out a trial at the point $x^{k+1}\in(x_{t-1},x_t)$, calculated using the following formulae
 \[
@@
 \]
-\[
+\begin{equation}\label{eq:142}
 x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N, \textrm{ if } 1<t<k+1.
-\]
+\end{equation}
 
@@
 Thus, at each iteration a new trial point is selected inside the interval, which has the greatest probability of finding the global minimum.
+
+Remark 2.
+In a different way, the global search algorithm has a geometric interpretation (see \cite{RefSergeyev2001}).
+Using trial points it is possible to design a support function (\textit{minorant}) for which characteristic $R(i)$ is the minimum value of the minorant in the interval $(x_{i-1},x_i), 1\leq i\leq k+1$, taken with the opposite sign.
+Then according to (\ref{eq:141}) new trial is carried out in the interval where the global minimizer of the current minorant is situated and the new trial point $x_{k+1}$ from (\ref{eq:142}) coincides with it.
+
 %new text end
 
-Remark 2.
+Remark 3.
 The proof of convergence of this algorithm is provided in \cite{RefStrongin2000}.
@@
 
-Let us illustrate the work of this \textit{global search algorithm} (GSA) using a dimension reduction for minimization of a two-dimensional multiextremal function
+Let us illustrate the work of this \textit{global search algorithm} (GSA) using a dimension reduction for minimization of a multiextremal function of two variables 
 \begin{eqnarray} \nonumber \label{eq:19}
@@
 The experiment used the method parameters $r=3$, $\epsilon=10^{-3}$, and the evolvent construction parameter $m=12$.
-Fig.~\ref{fig:1} shows the contour lines of a function generated according to (\ref{eq:19}) and the points of 1103 trials carried out using the method before the required accuracy was obtained.
+Fig.~\ref{fig:1} shows the level lines of a function generated according to (\ref{eq:19}) and the points of 1103 trials carried out using the method before the required accuracy was obtained.
 \begin{figure}
@@
   \includegraphics[width=0.75\textwidth]{fig2.jpg} 
-  \caption{Minimization of a two-dimensional function by the sequential algorithm}
+  \caption{Minimization of a function of two variables by the sequential algorithm}
   \label{fig:1}       % Give a unique label
@@
 Firstly, the part of computing process, in which the bulk of calculations is carried out, is parallelized.
-Secondly, the approach is applicable for a wide class of characteristical algorithms of multiextremal optimization \cite{RefGrishagin1997}.
+Secondly, the approach is applicable for a wide class of characteristical algorithms of multiextremal optimization \cite{RefSergeyev1994,RefSergeyevGrishagin,RefGrishagin1997,RefGrishagin2001}.
 %new text end
 
-Let us describe a change of rules of global search algorithm for the purpose of simultaneous (synchronous) carrying out in one iteration of method $p$ of trials.
-Let us make $k(n)$ the total number of trials, carried out after $n$ parallel iterations.
+Let us describe a change of rules of global search algorithm for the case when the algorithm carries out during one iteration $p>1$ trials simultaneously (each trial on a separate processor).
+Let $k(n)$ be the total number of trials, carried out after $n$ parallel iterations.
 
-Suppose $n\geq 1$  iterations of the method are carried out, in which trials were performed at $k=k(n)$ points $x^i,1\leq i \leq k$.
-Then points $x^{k+1},\dots,x^{k+p}$  of search trials of the next $n+1$ iteration are determined according to the rules.
+Suppose $n\geq 1$  iterations of the method have been carried out, in which trials were performed at $k=k(n)$ points $x^i,1\leq i \leq k$.
+Then points $x^{k+1},\dots,x^{k+p}$  of search trials of the next $n+1$ iteration are determined according to the following rules.
 
@@
 
-Step 5. Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), 1\leq j\leq p$, calculated using the formulas
+Step 5. Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), 1\leq j\leq p$, calculated using the formulae
 \[
@@
 
-The modifications of the parallel algorithm with the account of the information about the minimized function derivative are given in \cite{RefGergel1999}.
+The modifications of the parallel algorithm which takes into account the information about the minimized function derivative are given in \cite{RefGergel1999}.
 
@@
 With $p=4$ the number of iterations was reduced to $270$, and $1080$ trials were carried out.
-Fig.~\ref{fig:2} shows contour lines of the function with points of the trials performed by the method with $p=2$.
+Fig.~\ref{fig:2} shows level lines of the function with points of the trials performed by the method with $p=2$.
 Points of redundant trials carried out using a parallel algorithm are marked in the figure by "+".
@@
   \includegraphics[width=0.75\textwidth]{fig3.jpg} 
-  \caption{Minimization of a two-dimensional function by the parallel algorithm}
+  \caption{Minimization of a function of two variables by the parallel algorithm}
   \label{fig:2}
@@
 where $T(1)$ is the time required to solve a problem by a sequential algorithm, and $T(p)$ is the time for solving the same problem by a parallel algorithm in a system with $p$ computing elements.
-The characteristic of efficiency of a parallel algorithms (in relation to optimization algorithms) is also a speedup in iterations
+The characteristic of efficiency of parallel algorithms (in relation to optimization algorithms) is also a speedup in iterations
 \begin{equation}\label{eq:26}
@@
 \end{equation}
-where $n(1)$ is the number of trials carried out using the sequential method, and $n(p)$ is the number of trials carried out using the parallel method with $p$ processors.
+where $n(1)$ is the number of trials carried out by the sequential method, and $n(p)$ is the number of trials carried out by the parallel method with $p$ processors.
 This characteristic is especially important since for solving applied problems the time required to carry out a trial exceeds the time required to process its results.
@@
 Other trials points generally cannot coincide with the points generated by the sequential algorithm.
-The efficiency offered by the use of parallel processors can be reduced when such trials are performed.
+The efficiency owing to the use of parallel processors can be reduced when such trials are performed.
 Therefore, let us consider such trials as ``redundant'', and the value
@@
 	\item
-	The initial $q(l)$ trials of the sequential and parallel methods coincide, i.e.
+	The initial $q(l)$ trials of the sequential and parallel methods coincide, i.e.,
 	\[
@@
 
-Then a parallel algorithm of global search using two processors will be non-redundant (i.e. $s(2)=2$, $\lambda(2)=0$) while the following condition is satisfied 
+Then a parallel algorithm of global search using two processors will be non-redundant (i.e., $s(2)=2$, $\lambda(2)=0$) while the following condition is satisfied 
 \begin{equation}\label{eq:32}
@@
 
-Then, if the theorem conditions are satisfied, the parallel algorithm of global search with $Q+1$ processors will be non-redundant (i.e.
-$s(Q+1)= Q+1$, \mbox{$\lambda(Q+1)=0$}), while condition (\ref{eq:32}) is satisfied.
+Then, if the theorem conditions are satisfied, the parallel algorithm of global search with $Q+1$ processors will be non-redundant (i.e., $s(Q+1)= Q+1$, \mbox{$\lambda(Q+1)=0$}), while condition (\ref{eq:32}) is satisfied.
 
-The theorem corollary plays a special role in solving multidimensional problems reduced to one-dimensional problems by means of the Peano-like evolvent $y(x)$.
+The theorem corollary plays a special role in solving multidimensional problems reduced to univariate problems by means of the Peano-like evolvent $y(x)$.
 The evolvent $y(x)$, which is an approximation to the Peano curve, has the effect of ``splitting'' of a global minimum point $y^\ast\in D$ to several preimages in the interval $[0,1]$.
@@
 
-The parallel algorithm of global search has demonstrated a good speedup in implementation on CPU in systems with a small number of processors.
-In \cite{RefGrishagin1997,RefStrongin2003} a speedup of algorithm close to linear was experimentally verified when carrying out 2-4 parallel trials.
-The results obtained with the parallel GSA for minimization of 100 two-dimensional multiextremal functions described by formula (\ref{eq:19}) are shown in table \ref{tab:1}.
-All experiments used method parameters $r=2.9$, and $\epsilon=10^{-3}$, and the global minimum was found with the required accuracy in all $100$ problems.
-The table shows the number of used processors $p$, the average number of trials $n(p)$ rounded to integers and average speedup in iterations $s(p)$ from (\ref{eq:26}).
-
 \begin{table}
-	\caption{Results of minimization of $100$ two-dimensional functions}
+	\caption{Results of minimization of $100$ functions of two variables}
 	\label{tab:1}
@@
 
+The parallel algorithm of global search has demonstrated a good speedup in implementation on CPU in systems with a small number of processors.
+In \cite{RefGrishagin1997,RefStrongin2003} a speedup of algorithm close to linear was experimentally verified when carrying out 2-4 parallel trials.
+The results obtained with the parallel GSA for minimization of 100 multiextremal functions of two variables described by formula (\ref{eq:19}) are shown in table \ref{tab:1}.
+All experiments used method parameters $r=2.9$, and $\epsilon=10^{-3}$, and the global minimum was found with the required accuracy in all $100$ problems.
+The table shows the number of used processors $p$, the average number of trials $n(p)$ rounded to integers and average speedup in iterations $s(p)$ from (\ref{eq:26}).
+
 \section{Implementation on GPU} \label{sec:5}
@@
 
-With respect to methods of global optimization, parallel computing of many values of the objective function simultaneously is an operation which can be efficiently implemented on GPU.
-In the function value calculation procedure implemented on GPU transfers of information from CPU to GPU will be minimal: it is necessary only to transfer the coordinates of trial points to GPU and to receive the values of the function in these points.
-The functions that determine the processing of trial results according to algorithm and require working with a large volume of collected search information can be efficiently implemented on CPU.
+With respect to methods of global optimization, parallel computing many values of the objective function simultaneously is an operation which can be efficiently implemented on GPU.
+In the function value calculation procedure implemented on GPU transfers of information from CPU to GPU will be minimal: it is necessary only to transfer the coordinates of trial points to GPU and to receive the values of the function at these points.
+The routines that determine the processing of trial results according to algorithm and require working with a large volume of collected search information can be efficiently implemented on CPU.
 
@@
 
-The use of several hundred of parallel computing elements (as in modern GPU) might not give a speedup of the search process by hundreds of times.
-In this case, the conditions of the theorem of non-redundant parallelization can be violated: the number of local extremums will be less than the number of computing cores.
+The use of several hundred parallel computing elements (as in modern GPU) might not give a speedup of the search process by hundreds of times.
+In this case, the conditions of the theorem of non-redundant parallelization can be violated: the number of local extrema will be less than the number of computing cores.
 Then (according to the theorem and its corollary) a parallel GSA will generate redundant trial points.
 Nevertheless, despite certain redundancy, the overall operating time of a GPU algorithm will be less than the operating time of a CPU algorithm.
-This has been verified by computing experiments, the results of which are given in section \ref{sec:6}.
+This has been verified by computational experiments, the results of which are given in section \ref{sec:6}.
 
@@
 
-Computing experiments were carried out on one of the nodes of a high-performance cluster of the State University of Nizhny Novgorod.
+Computational experiments were carried out on one of the nodes of a high-performance cluster of the State University of Nizhny Novgorod.
 The cluster node includes 2 Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM, and NVIDIA Tesla M2090 GPU.
-The CPU has 8 cores (i.e.
-16 physical cores and 16 virtual cores in hyper-threading mode are available in the node).
+The CPU has 8 cores (i.e., 16 physical cores and 16 virtual cores in hyper-threading mode are available in the node).
 The graphics processing unit includes 16 multiprocessors (512 cores).
@@
 
-Therefore, for the purpose of imitation of the computational complexity inherent to applied problems of optimization \cite{RefBarkalov2013}, calculation of the objective function in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (series summation from 20 thousand elements).
+Therefore, for the purpose of simulation of the computational complexity inherent to applied problems of optimization \cite{RefBarkalov2013}, calculation of the objective function in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (series summation from 20 thousand elements).
 
@@
 
-As is seen from this table, there is a speedup of the algorithm in the case of an increase in the number of trials at one iteration up to the number of cores on GPU, with subsequent saturation.
+As it seen from this table, there is a speedup of the algorithm in the case of an increase in the number of trials at one iteration up to the number of cores on GPU, with subsequent saturation.
 Owing to a weak information dependence the size of the thread block $p_b$ does not have decisive effect, therefore let us use $p_b=64$ in all subsequent experiments.
@@
 
-After the parameters of the GPU algorithm have been determined let us make the main computing experiments on a series of multidimensional multiextremal problems.
+After the parameters of the GPU algorithm have been determined let us make the main computational experiments on a series of multidimensional multiextremal problems.
 In work \cite{RefGaviano} a GKLS generator allowing generation of multiextremal optimization problems with known properties (number of local minima, size of their domains of attraction, global minimum point, etc.) is described.
 
-The results of numerical comparison of the three sequential algorithms -- DIRECT \cite{RefJones}, DIRECT\textit{l} \cite{RefGablonsky} and global search algorithm (GSA) from section \ref{sec:2} -- are provided below (results of the work of the first two algorithms are given in paper \cite{RefGaviano}).
+The results of numerical comparison of three sequential algorithms -- DIRECT \cite{RefJones}, DIRECT\textit{l} \cite{RefGablonsky} and global search algorithm (GSA) from section \ref{sec:2} -- are provided below (results of the work of the first two algorithms are given in paper \cite{RefGaviano}).
 Numerical comparison was carried out on function classes \textit{Simple} and \textit{Hard} of dimension 4 and 5 from \cite{RefGaviano} since solving a problems of dimension 2 and 3 demands a small number of iterations and it is impractical to use GPU for solving these problems.
-Global minimum $y^\ast$ was considered to be found, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e.
-$\left\|y^k-y^\ast\right\|\leq\delta$.
+Global minimum $y^\ast$ was considered to be found, if the algorithm generated trial point $y^k$ in $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^\ast\right\|\leq\delta$.
 The size of the vicinity was selected (according to \cite{RefGaviano}) as $\delta = \left\|b-a\right\|\sqrt[N]{\Delta}$, $N$ -- problem dimension, $a$ and $b$ -- borders of search domain $D$, parameter $\Delta=10^{-6}$ at $N=4$ and $\Delta=10^{-7}$ at $N=5$.
@@
 Table \ref{tab:3} shows that GSA outperforms DIRECT and DIRECT\textit{l} methods on all classes of problems in terms of average number of iterations.
-In class 5-\textit{Hard} all the three methods failed to solve all problems: DIRECT did not solve 16 problems, DIRECT\textit{l} and GSA, 4 problems.
+In class 5-\textit{Hard} all three methods failed to solve all problems: DIRECT did not solve 16 problems, DIRECT\textit{l} and GSA, 4 problems.
 
@@
 
-The data show five-fold time speedup, when using 32 cores for all classes (except 5-\textit{Hard}).
-The lack of expected speedup in class 5-\textit{Hard} in transition from 16 to 32 cores is explained by an increase in the average number of iterations and by the fact that out of 32 cores 16 are physical, and other 16 are virtual in hyper-threading mode, i.e.
-a real gain in performance in transition from 16 to 32 cores is less than twofold.
+The data show fivefold time speedup, when using 32 cores for all classes (except 5-\textit{Hard}).
+The lack of expected speedup in class 5-\textit{Hard} in transition from 16 to 32 cores is explained by an increase in the average number of iterations and by the fact that of 32 cores 16 are physical, and other 16 are virtual in hyper-threading mode, i.e., a real gain in performance in transition from 16 to 32 cores is less than twofold.
 
@@
 The number of the used threads $p$ on GPU varied, all other parameters of the method did not change.
-The number of threads, which is greater, than the number of cores, was chosen to show the following effect: the number of iterations executed by the algorithm will be reduced in case of an increase of the threads number, while the algorithm operation time will not be reduced.
+The number of threads, which is greater than the number of cores, was chosen to show the following effect: the number of iterations executed by the algorithm will be reduced in case of an increase of the threads number, while the algorithm operation time will not be reduced.
 The minimum time of operation is reached, if the number of used threads does not surpass the number of GPU cores.
@@
 The results of our experiments show some obvious advantages of the GPU algorithm in comparison with the CPU algorithm.
-The algorithm using the maximum quantity of cores of CPU works slower than any of the given GPU implementations.
+The algorithm, which uses the maximal number of cores of CPU, works slower than any of the given GPU implementations.
 In addition, the results show a considerable speedup in the number of iterations.
@@
 
-For the purpose of experimental demonstration of theoretical properties of the considered parallel algorithms, computing experiments on a series of several hundred test problems of different dimensions were performed.
-The results obtained with the sequential global search algorithm show its superiority over other known methods of global optimization.
+For the purpose of experimental demonstration of theoretical properties of the considered parallel algorithms, computational experiments on a series of several hundred test problems of different dimensions were performed.
+The results obtained with the sequential global search algorithm show its superiority over several known methods of global optimization.
 At the same time the parallel algorithm shows a good speedup and low redundancy (both GPU and CPU version), and the GPU version of the parallel algorithm shows a speedup of 2 to 6 times in comparison with the fastest parallel CPU version.
@@
 
-The authors of the paper express their gratitude to Ilya Lebedev for assistance in carrying out the computing experiments.
+The authors of the paper express their gratitude to Ilya Lebedev for assistance in carrying out the computational experiments.
 
@@
 
+\bibitem{RefKindratenko}
+Kindratenko, V. (ed.): Numerical Computations with GPUs. Springer (2014).
+
 \bibitem{RefStrongin2000}
 Strongin, R.G., Sergeyev, Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2000)
+
+\bibitem{RefLebedevBarkalov}
+Lebedev, I.G., Barkalov, K.A.: A GPU implementation of a parallel global search algorithm. PNRPU Aerospace Engineering Bulletin 36, 64--82 (2014) (in Russian)
+
+\bibitem{RefSergeyev2006}
+Sergeyev Ya.D., Kvasov D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim. 16(3), 910--937 (2006)
+
+\bibitem{RefZilinskas}
+Zilinskas, J.: Branch and bound with simplicial partitions for global optimization. Mathematical Modelling and Analysis 13(1), 145--159 (2008)
+
+\bibitem{RefSergeyev2013}
+Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013)
+
+\bibitem{RefSergeyev2001}
+Molinaro, A., Pizzuti, C., and Sergeyev, Ya.D.: Acceleration tools for diagonal information global optimization algorithms. Computational Optimization and Applications 18, 5--26 (2001)
 
@@
 
-\bibitem{RefGergel1999}
-Gergel, V.P., Sergeyev, Ya.D.: Sequential and parallel algorithms for global minimizing functions with lipschitzian derivatives. Computers and Mathematics with Applications, 37(4-5), 163--179 (1999)
-
 \bibitem{RefSergeyev1994}
@@
 
-\bibitem{RefSergeyev2006}
-Sergeyev Ya.D., Kvasov D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim. 16(3), 910--937 (2006)
-
-\bibitem{RefZilinskas}
-Zilinskas, J.: Branch and bound with simplicial partitions for global optimization. Mathematical Modelling and Analysis 13(1), 145--159 (2008)
+\bibitem{RefSergeyevGrishagin}
+Sergeyev, Ya.D., Grishagin, V.A.:  Sequential and parallel global optimization algorithms. Optimization Methods and Software. 3, 111--124 (1994)
 
@@
 
+\bibitem{RefGrishagin2001}
+Sergeyev Ya.D, Grishagin V.A.: Parallel asynchronous global search and the nested optimization scheme. Journal of Computational Analysis and Applications. 3(2), 123--145 (2001)
+
+\bibitem{RefGergel1999}
+Gergel, V.P., Sergeyev, Ya.D.: Sequential and parallel algorithms for global minimizing functions with Lipschitzian derivatives. Computers and Mathematics with Applications, 37(4-5), 163--179 (1999)
+
 \bibitem{RefStrongin2003}
 Strongin, R.G., Sergeyev, Ya.D.: Global optimization: fractal approach and non-redundant parallelism. J. Glob. Optim. 27(1), 25--50 (2003)
-
-\bibitem{RefSergeyev2013}
-Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013)
 
@@
 
+%10.	V. Gergel, V. Grishagin, R. Israfilov, Local Tuning in Nested Scheme of Global Optimization, Procedia Computer Science 51, pp. 865–874, 2015.
+
 \end{thebibliography}

### 2019-10-07 2016/JoGO_GPU/BarkalovGergel.tex
JOGO2016 paper final version

@@
 For example, the objective function can be defined by solving SLAE, ODE, PDE, or Monte Carlo simulation.
-Currently, GPU may be used to solve all these types of problems \cite{RefKindratenko}.
-Therefore, calculation of values of an optimized function can be done using a GPU, and the main objective of the algorithm (working at CPU) is to efficiently select points of parallel trials.
+Currently, GPU can be used to solve all these types of problems.
+%rev4 text begin
+Moreover, a set of problems can be solved on GPU simultaneously.
+For example, algorithm of solving of a large amount of independent SLAE on GPU is considered in \cite{RefKindratenko}, chapter 5.
+Results of simultaneously solving $2 \cdot 10^4$ systems of variable size are presented.
+Algorithm of parallel integration of large numbers of independent ODE systems on GPU is discussed in \cite{RefKindratenko}, chapter 8.
+Experimental results that show good speedup in solving a set (from $10^3$ up to $10^5$) of independent ODE systems are presented.
+Assuming that the objective function in global optimization problem is given as SLAE or ODE system many function values can be computed simultaneously.
+%rev4 text end
+
+Therefore, calculation of an objective function values can be done using a GPU, and the main goal of the algorithm	 (working at CPU) is to efficiently select points of parallel trials.
 The article contains the results of an investigation into the GPU implementation of the global search parallel algorithm developed according to the information-statistical approach presented in the monograph \cite{RefStrongin2000}.
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \begin{minipage}{0.32\linewidth}
@@
 \end{minipage}
-\hfill
 \begin{minipage}{0.32\linewidth}
@@
 \end{minipage}
-\caption{Elvolvents in two dimensions with (a) $m=3$, (b) $m=4$ and (c) $m=5$}
+\caption{Evolvents in two dimensions with (a) $m=3$, (b) $m=4$ and (c) $m=5$}
 \label{fig:0}
@@
 
+\begin{figure}
+	\center
+  \includegraphics[width=0.75\textwidth]{fig3.jpg} 
+  \caption{Minimization of a function of two variables by the parallel algorithm}
+  \label{fig:2}
+\end{figure}
+
 \section{Synchronous parallel algorithm} \label{sec:3}
@@
 The concept of redundancy is discussed in more detail in the following section.
-\begin{figure}
-	\center
-  \includegraphics[width=0.75\textwidth]{fig3.jpg} 
-  \caption{Minimization of a function of two variables by the parallel algorithm}
-  \label{fig:2}
-\end{figure}
 
@@
 Usually, such a calculation is nothing else than a summation of several (according to problem dimension) values of elementary functions.
-For example, the well known Rastrigin test function, the global minimum of which is located in $y^\ast = (0,\dots,0)$ with $\varphi(y^\ast)=0$, is set by the expression
+For example, the well known Rastrigin test function
 \begin{equation}\label{eq:33}
-\varphi(y)=10N+\sum^{N}_{i=1}\left[y^2_i-10\cos(2\pi y_i)\right], \; -2.2\leq y_i \leq 1.8, \; 1\leq i \leq N.
+\varphi(y)=10N+\sum^{N}_{i=1}\left[y^2_i-10\cos(2\pi y_i)\right], \; -2.2\leq y_i \leq 1.8, \; 1\leq i \leq N,
 \end{equation}
+has the global minimum value $\varphi(y^\ast)=0$ at the point $y^\ast = (0,\dots,0)$.
 
@@
 The algorithm, which uses the maximal number of cores of CPU, works slower than any of the given GPU implementations.
-In addition, the results show a considerable speedup in the number of iterations.
+In addition, the results show a considerable decrease of the number of iterations.
 For example, on average only 353 parallel iterations on GPU were sufficient to solve problems of the most difficult class considered -- 5-\textit{Hard}, whereas even with the use of all computing CPU cores over eleven thousand iterations were required.
@@
 \end{document}
-% end of file template.tex

### 2019-11-01 2020/PAVT_Nested_Full/pavt_paper.tex
Templates for PAVT2020 papers

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{svproc}
%
% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\usepackage{graphicx}
\usepackage{marvosym}

\usepackage[utf8]{inputenc}
\usepackage[english, russian]{babel} % Русские и английские переносы

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\usepackage{hyperref}
\def\UrlFont{\rmfamily}

\def\orcidID#1{\unskip$^{[#1]}$}
\def\letter{$^{\textrm{(\Letter)}}$}

\begin{document}
\mainmatter              % start of a contribution
%
\title{ Combining local and global methods in a parallel nested optimization scheme 
\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
}
%
\titlerunning{ABBR}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Konstantin Barkalov\letter\orcidID{0000-0001-5273-2471} \and Ilya Lebedev\orcidID{0000-0002-8736-0652} \and Maria Kocheganova \orcidID{0000-0002-4722-6299}}
%
\authorrunning{Konstantin Barkalov et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Konstantin Barkalov and Ilya Lebedev, and Maria Kocheganova}
%
\institute{Lobachevsky State University of Nizhni Novgorod, Russia  \\
	\email{konstantin.barkalov@itmm.unn.ru}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}

В статье рассматриваются задачи глобальной оптимизации и численные методы их решения. О характере зависимости целевой функции от ее параметров делается предположение, что она является многоэкстремальной лишь по некоторым из переменных, а зависимость от остальных параметров носит локальный характер. Задачи такого типа могут возникать при идентификации математических моделей по результатам экспериментов. Предложена параллельная схема вычислений, которая учитывает данную особенность. Новая схема основана на идее рекурсивной оптимизации, когда на верхнем уровне рекурсии проводится оптимизации по параметрам, влияющим глобально (с помощью алгоритма глобального поиска), а на нижнем -- решение задач локальной оптимизации (с помощью локальных методов). Возникающие при этом локальные подзадачи не будут оказывать влияния друг на друга, и их решение можно проводить параллельно. Проведены численные эксперименты на нескольких сотнях тестовых задач, подтверждающие эффективность предложенной схемы параллельных вычислений.

\keywords{Global optimization $\cdot$ Local optimization $\cdot$ Recursive optimization scheme $\cdot$ Parallel algorithms}
\end{abstract}

\section{Introduction}

%Вводные слова

Примером задачи, в которой наблюдается разный характер зависимостей от параметров, может являться следующая задача аппроксимации.

Пусть в ходе эксперимента получены $m$ значений $u_j = u(t_j), 1 \leq j \leq m, $ функции $u(t)$, причем известно, что аналитическое выражение для  $u(t)$ имеет вид
\begin{equation}\label{ex_func}
u(t) = u(t,\omega,c)=\sum^{n}_{i=0}\left[c_{2i+1}\sin(\omega_it) + c_{2i+2}\cos(\omega_it)\right].	
\end{equation}
В данном выражении $c=\left(c_1,\ldots,c_{2n+2}\right)$, $\omega=\left(\omega_0,\omega_1,\ldots,\omega_n\right)$ являются неизвестными параметрами, определяющими конкретную функцию $u(t)$.

Введем меру отклонения функции $u(t,\omega,c)$ от экспериментальных данных как сумму квадратов
\[
\Delta(\omega,c)= \sum^{m}_{j=1}\left[u_j-u(\tau_j,\omega,c)\right]^2. 
\]
Now, following the idea of least squares fitting, it is possible to present the approximation problem as the problem of minimizing the objective function
\begin{equation}\label{ex_prob}
\Delta(\omega^*,c^*) = \min_{\omega,c} \Delta(\omega,c).	
\end{equation}
Решение $(\omega^*,c^*)$ данной задачи будет определять наилучшее приближение.

Задача (\ref{ex_prob}) может быть записана в рекурсивной форме
\begin{equation}\label{outer_prob}
\varphi(\omega^*) = \min_\omega \varphi{\omega}
\end{equation}
\begin{equation}\label{inner_prob}
\varphi(\omega) = \min_c \Delta{\omega,c}.
\end{equation}
Вложенная подзадача (\ref{inner_prob}) является классической линейной задачей наименьших квадратов, и ее решение can be obtained by solving a system of linear algebraic equations regarding the unknown $c$, which can be done, e.g., by Gaussian elimination. Внешняя задача (\ref{outer_prob}) будет являться, вообще говоря, многоэкстремальной, и для ее решения нужно применять методы глобальной оптимизации.

%Структура статьи

\section{Nested optimization scheme}

%Постановка задачи в целом, условие Липшица, многошаговая схема с 2 уровнями
%Внешний уровень - глобальная задача, вложенный уровень - локальная задача

\section{Parallel global search algorithm}

%Постановка задачи глобальной оптмизации
%Редукция размерности - Пеано
%Параллельный алгоритм

\section{Local search algorithms}

%Локальные задачи и локальные методы
%Стандартные библиотеки - будем использовать одну из них

\section{Numerical experiments}

%Сравнение разных локальных методов
%Эффективность распараллеливания

\section{Conclusion}

%
% ---- Bibliography ----
%
\bibliographystyle{spmpsci}
\bibliography{bibliography}{}

\end{document}

### 2019-11-01 2020/PAVT_Poster/poster-ru.tex
Templates for PAVT2020 papers

\documentclass[11pt, oneside, a4paper]{article}
\usepackage[utf8]{inputenc}
%\usepackage[cp1251]{inputenc} % кодировка
\usepackage[english, russian]{babel} % Русские и английские переносы
\usepackage{graphicx}          % для включения графических изображений
\usepackage{cite}              % для корректного оформления литературы
\usepackage{pavt-ru}

\begin{document}

% \title - название статьи
% \authors - список авторов

\title{Параллельный алгоритм минимизации многоэкстремальных функций, имеющих разрывы \footnote{Исследование выполнено за счет гранта Российского научного фонда (проект № 16-11-10150).}}

\authors{К.А.~Баркалов, М.А.~Усова}
\organizations{Нижегородский государственный университет им. Н.И. Лобачевского}

Описание плаката (с текстом в объеме \emph{одной страницы формата А4}) содержит информацию о планах и начальных результатах недавно начатого научного исследования. Описание плаката, в отличие от полных и коротких статей, \emph{не имеет аннотации и списка ключевых слов, а также переводов названия, списка авторов и др. элементов на английский язык}. В то же время описание плаката может содержать рисунки, таблицы и список литературы. Страница статьи, представляемой на конференцию, должна иметь \emph{размеры}
$297 \times 210$~мм. Все \emph{поля} страницы должны иметь одинаковый размер~--- 25~мм. Не допускается использование нумерации, принудительных разрывов страниц и колонтитулов.

\emph{Название} оформляется шрифтом размером 16~пт с выравниванием по центру страницы. Сверху от остального текста название отделяется одной пустой строкой.

\emph{Список авторов} оформляется шрифтом размером 12~пт с выравниванием по центру и отделяется от названия одной пустой строкой размером 16~пт. Авторы перечисляются через запятую, фамилия ставится справа от инициалов. \emph{Организация} оформляется шрифтом размером 12~пт с выравниванием по центру и отделяется от списка авторов одной пустой строкой размером 6~пт. В качестве организации необходимо указать полное наименование организации, в которой работают авторы. В случае, если авторы из разных организаций, через запятую указывается две или более организаций, и принадлежность автора к соответствующей организации обозначается при помощи сносок.

В тексте описания плаката используется \emph{шрифт} Times New Roman. \emph{Абзац} оформляется шрифтом размером 11~пт с выравниванием по ширине страницы, одинарным интервалом между строками и автоматической расстановкой переносов. Абзацы не разделяются интервалами и начинаются с красной строки с отступом 7~мм. Все \emph{рисунки и таблицы} должны иметь подпись c шрифтом размером 10~пт, отступ сверху и снизу 6 пт и выравнивание по центру страницы. Подпись к таблице помещается над таблицей, подпись к рисунку~--- под рисунком.

\emph{Благодарности и ссылки на грант} оформляются в виде сноски к названию статьи и обозначаются символом *. \emph{Перекрестные ссылки} на литературу заключаются в квадратные скобки и перечисляются через запятую или тире, например \cite{eremin}, \cite{eremin,sokolinsky}, \cite{eremin,sokolinsky,stonebraker}. \emph{Библиографические источники} оформляются в виде нумерованного арабскими цифрами списка шрифтом размером 11~пт с выравниванием по левому краю.

\begin{biblio}

\bibitem{eremin}
Ерёмин~И.И. Фейеровские методы для задач выпуклой и линейной оптимизации. 
Челябинск: Изд-во ЮУрГУ, 2009. 200~с.

\bibitem{sokolinsky}
Соколинский~Л.Б. Организация параллельного выполнения запросов в многопроцессорной 
машине баз данных с иерархической архитектурой // Программирование. 2001. №~6. С.~13--29.

\bibitem{stonebraker}
Stonebraker~M., Kemnitz~G. The POSTGRES Next-generation Database Management System // Communications of the ACM. 1991. Vol.~34, No.~10. P.~78--92.
DOI:~\href{http://dx.doi.org/10.1145/125223.125262}{10.1145/125223.125262}.

\end{biblio}
\end{document}

### 2019-11-01 2020/PAVT_Nested_Full/pavt_paper.tex
A half og introduction

@@
 
+Новым элементом, исследуемым в данной работе, является предположение о разном характере зависимости целевой функции от разных переменных параметров. 
+Предполагается, что целевая функция является многоэкстремальной лишь по части переменных. Остальные переменные влияют локально, т.е. по ним функция является одноэкстремальной.
+
+В этом случае решение задачи можно организовать по схеме вложенной (рекурсивной) оптимизации.
+Решение  первой (многоэкстремальной) подзадачи, для которой требуется использовать сложные алгоритмы глобальной оптимизации, будет проводиться на верхнем уровне рекурсии.
+Одноэкстремальные подзадачи (каждая из которых соответствует фиксированному набору значений первой части параметров) будут решаться на нижнем уровне.
+Для решения одноэкстремальных задач можно применять эффективные методы выпуклой оптимизации.
+
 Примером задачи, в которой наблюдается разный характер зависимостей от параметров, может являться следующая задача аппроксимации.
@@
 \begin{equation}\label{outer_prob}
-\varphi(\omega^*) = \min_\omega \varphi{\omega}
+\varphi(\omega^*) = \min_\omega \varphi(\omega)
 \end{equation}
 \begin{equation}\label{inner_prob}
-\varphi(\omega) = \min_c \Delta{\omega,c}.
+\varphi(\omega) = \min_c \Delta(\omega,c).
 \end{equation}
@@
 \section{Nested optimization scheme}
+
+Рассмотрим задачу оптимизации вида
+\begin{eqnarray}\label{main_problem}
+& f(x^\ast,y^\ast)=\min{\left\{\varphi(x,y):x\in S, y\in D\right\}}, \nonumber \\
+& S=\left\{x\in R^M: a_i\leq x_i \leq b_i, 1\leq i \leq M\right\}, \nonumber \\
+& D=\left\{y\in R^N: c_i\leq y_i \leq d_i, 1\leq i \leq N\right\}. \nonumber
+\end{eqnarray}
+
+Будем предполагать, что при любом фиксированном наборе значений $x'$ функция $f(x',y)$ является многоэкстремальной и удовлетворяет условию Липшица по переменной $y$, т.е.
+% условие Липшица. 
+Одновременно с этим при любом фиксированном наборе значений $y'$  функция $f(x,y')$ является унимодальной, т.е. функция $f(x,y')$ имеет единственную точку минимума $x^*$.
+
+Учет подобной особенности рассматриваемого класса задач может существенно снизить вычислительную сложность поиска оптимума.
+В самом деле, в соответствии с известной схемой рекурсивной оптимизации (ссылка) решение исходной задачи можно свести к задаче поиска глобального минимума функции $\varphi(y)$
+\[
+\varphi(y^*) = \min_{y\in D}  \varphi (y)
+%\min \left\{ f(x,y): x\in S, y \in D \right\} = \min_{y\in D}  \left\{ \min_{x\in S} f(x,y) \right\}.
+\]
+где 
+\[
+\varphi(y) = \min_{x\in S} f(x,y).
+\]
+В соответствии с данной схемой вычисление одного значения функции $\varphi (y)$ подразумевает решение локальной задачи , которое может быть выполнено одним из методов выпуклой оптимизации.
 

### 2019-11-05 2020/PAVT_Nested_Full/pavt_paper.tex
First version in Russian

--- 2019-11-01 2020/PAVT_Nested_Full/pavt_paper.tex
@@
 \usepackage{marvosym}
+
+%\pdfmapfile{+txfonts.map}
 
@@
 
-%Вводные слова
+Задачи глобальной оптимизации часто возникают в различных областях современной науки и техники.
+Например, в биологии и химии актуальны проблемы поиска конфигураций различных химических соединений, соответствующих минимуму энергии взаимодействия.
+В частности, такие задачи возникают при разработке новых лекарственных препаратов \cite{Sulimov}.
 
-Примером задачи, в которой наблюдается разный характер зависимостей от параметров, может являться следующая задача аппроксимации.
+Классическим стало применение методов глобальной оптимизации для идентификации параметров моделей по данным экспериментов.
+В задачах такого вида требуется провести поиск значений неизвестных параметров модели, при которых результаты расчетов по модели  близки к результатам, полученным экспериментально.
+
+Число параметров, которые требуется идентифицировать подобным образом, для сложных математических моделей может составлять десятки и сотни \cite{Nurislamova2016,Akhmadullina2017}.
+В задачах подобной размерности регулярные методы поиска глобального решения не могут быть применены из-за чрезвычайно больших вычислительных затрат на покрытие области точками испытаний.
+Это остается справедливым даже в случае использования эффективных параллельных алгоритмов, строящих существенно неравномерные покрытия области \cite{Paulavicius2011,Evtushenko2009,Jones2009}.
+
+Однако в таких задачах часто можно указать параметры, которые будут оказывать наибольшее влияние на результат.
+Число таких параметров, как правило, является небольшим.
+Остальные параметры либо оказывают незначительное влияние (в этом случае многоэкстремальность можно не учитывать), либо зависимость от второй группы параметров носит локальный характер.
+
+Таким образом, актуальной является разработка параллельных алгоритмов минимизации существенно многомерных функций (сотни переменных), в которых учитывается разный характер зависимости целевой функции от разных групп параметров.
+А именно, предполагается, что целевая функция является многоэкстремальной лишь по части переменных.
+Остальные переменные влияют локально, т.е.
+по ним функция является одноэкстремальной.
+В этом случае решение задачи можно организовать по схеме вложенной (рекурсивной) оптимизации.
+Решение  многоэкстремальной подзадачи, для которой требуется использовать сложные алгоритмы глобальной оптимизации, будет проводиться на верхнем уровне рекурсии.
+Одноэкстремальные подзадачи (каждая из которых соответствует фиксированному набору значений первой части параметров) будут решаться на нижнем уровне.
+Для решения одноэкстремальных задач можно применять эффективные методы локальной оптимизации, линейного программирования или же линейной алгебры (в зависимости от характера локальных подзадач).
+
+Текст статьи, отражающей результаты проведенного исследования, построен следующим образом. 
+В Section 2 дано описание схемы вложенной оптимизации с учетом разных свойств функции на разных уровнях рекурсии.
+Приведен конкретный пример задачи указанного класса.
+Описана общая схема организации параллельных вычислений.
+Section 3 посвящен описанию параллельного алгоритма глобального поиска с использованием с использованием space-filling curves. Приведена формулировка параллельного алгоритма, пример его работы.
+Section 4 обсуждаются методы решения локальных подзадач в рамках схемы рекурсивной оптимизации. Дается кратное описание свойств задач и известных методов их решения. 
+Section 5 содержит результаты численных экспериментов. Здесь производится оценка масшабируемости предложенного параллельного алгоритма при решении серии тестовых задач.  
+Section 6 concludes the paper.
+
+\section{Nested optimization scheme}
+
+Рассмотрим задачу оптимизации вида
+\begin{eqnarray}\label{main_problem}
+& f(x^\ast,y^\ast)=\min{\left\{f(x,y):x\in S, y\in D\right\}}, \nonumber \\
+& S=\left\{x\in R^M: a_i\leq x_i \leq b_i, 1\leq i \leq M\right\}, \nonumber \\
+& D=\left\{y\in R^N: c_i\leq y_i \leq d_i, 1\leq i \leq N\right\}. \nonumber
+\end{eqnarray}
+
+Будем предполагать, что при любом фиксированном наборе значений $\overline{x}$ функция $f(\overline{x},y)$ является многоэкстремальной и удовлетворяет условию Липшица
+\[
+\left|f(\overline{x},y')-f(\overline{x},y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in D,\; 0<L<\infty,
+\]
+with the constant $L$ unknown a priori.
+Одновременно с этим при любом фиксированном наборе значений $\overline{y}$  функция $f(x,\overline{y})$ является унимодальной, т.е. функция $f(x,\overline{y})$ имеет единственную точку минимума $x^*$.
+
+Учет подобной особенности рассматриваемого класса задач может существенно снизить вычислительную сложность поиска оптимума.
+В самом деле, в соответствии с известной схемой рекурсивной оптимизации (ссылка) решение исходной задачи можно свести к задаче поиска глобального минимума функции $\varphi(y)$
+\begin{equation}\label{global_problem}
+\varphi(y^*) = \min_{y\in D}  \varphi (y)
+%\min \left\{ f(x,y): x\in S, y \in D \right\} = \min_{y\in D}  \left\{ \min_{x\in S} f(x,y) \right\}.
+\end{equation}
+где 
+\begin{equation}\label{local_problem}
+\varphi(y) = \min_{x\in S} f(x,y).
+\end{equation}
+В соответствии с данной схемой вычисление одного значения функции $\varphi (y)$ (данный процесс для задачи \ref{global_problem} будем называть search trial) подразумевает решение унимодальной задачи (\ref{local_problem}), которое может быть выполнено одним из методов локальной оптимизации.
+Эффективность и проработанность методов локальной оптимизации позволяет решать подзадачи (\ref{local_problem}) с высокой точностью.
+Соответственно, можно рассматривать задачу (\ref{global_problem}) как задачу глобальной оптимизации, в которой значение функции может быть вычислено точно, но данная операция является трудоемкой.
+
+Конкретным примером задачи, в которой наблюдается разный характер зависимостей от разных групп параметров, может являться следующая задача аппроксимации.
 
@@
 \begin{equation}\label{outer_prob}
-\varphi(\omega^*) = \min_\omega \varphi{\omega}
+\varphi(\omega^*) = \min_\omega \varphi(\omega)
 \end{equation}
 \begin{equation}\label{inner_prob}
-\varphi(\omega) = \min_c \Delta{\omega,c}.
+\varphi(\omega) = \min_c \Delta(\omega,c).
 \end{equation}
 Вложенная подзадача (\ref{inner_prob}) является классической линейной задачей наименьших квадратов, и ее решение can be obtained by solving a system of linear algebraic equations regarding the unknown $c$, which can be done, e.g., by Gaussian elimination.
-Внешняя задача (\ref{outer_prob}) будет являться, вообще говоря, многоэкстремальной, и для ее решения нужно применять методы глобальной оптимизации.
+Одновременно с этим внешняя задача (\ref{outer_prob}) будет являться, вообще говоря, многоэкстремальной.
 
-%Структура статьи
-
-\section{Nested optimization scheme}
-
-%Постановка задачи в целом, условие Липшица, многошаговая схема с 2 уровнями
-%Внешний уровень - глобальная задача, вложенный уровень - локальная задача
+Таким образом, для решения задачи (\ref{main_problem}) в рекурсивной схеме (\ref{global_problem}), (\ref{local_problem}) можно применять параллельный алгоритм глобального поиска для решения внешних подзадач (\ref{global_problem}).
+При этом на каждой итерации алгоритма будет порождаться множество независимых подзадач (\ref{local_problem}), решение которых может проводиться параллельно с помощью локальных методов.
+Общая схема организации вычислений с использованием нескольких процессов приведена на Fig.~\ref{MPI_tree}. 
+% тут нужен наш традиционный рисунок - дерево процессов
+Процессы параллельной программы будут образовывать дерево. 
+Корневой процесс будет решать задачу (\ref{global_problem}) и распределять подзадачи (\ref{local_problem}) между процессами-потомками.
+Каждая подзадача решается в отдельном процессе, обмен данными между процессами может быть организован с помощью MPI.
+Пересылки данных между процессами будут минимальные: требуется лишь передать от корня к потомкам координаты точек испытаний, и получить обратно значения функции в этих точках.
+Подробное обсуждение алгоритмов, используемых на разных уровнях дерева процессов, приведено в следующих разделах.
 
@@
 
-%Постановка задачи глобальной оптмизации
-%Редукция размерности - Пеано
-%Параллельный алгоритм
+В ННГУ им.
+Н.И.
+Лобачевского был предложен \cite{Barkalov2018,Strongin2018,globalizerSystem} эффективный параллельный алгоритм глобальной оптимизации, предназначенный для решения задач вида (\ref{global_problem}) .
+Основная идея в распараллеливании состоит в такой организации вычислений, при которой параллельно проводятся несколько испытаний.
+Этот подход характеризуется эффективностью (распараллеливается именно та часть вычислительного процесса, в котором выполняется основной объем вычислений) и общностью (применим для широкого класса алгоритмов глобальной оптимизации).
+
+В разработанных методах используется оригинальный прием снижения размерности решаемой задачи. 
+Редукция размерности основана на фундаментальном факте, согласно которому $N$-мерный гиперпараллелепипед $D$ и отрезок $[0,1]$ вещественной оси являются равномощными множествами, и отрезок $[0,1]$ может быть непрерывно отображен в $D$ с помощью Peano curve $y(x)$, i.e.
+$D = \left\{y(x):0\leq x\leq 1\right\}$.
+
+Используя этот факт, можно свести решение задачи минимизации многомерной функции $\varphi(y)$ к минимизации одномерной функции $\varphi(y(x))$
+\begin{equation}\label{1d_problem}
+\varphi(y (x^*)) = \min \left\{ \varphi (y(x)) : x\in [0,1] \right\}.
+\end{equation}
+
+The algorithms for the numerical construction of the Peano curve approximations (called evolvents)
+are given in \cite{Strongin2000,Sergeyev2013}.
+As an illustration, two evolvents are shown in Fig.~\ref{fig_evolvent}.
+Figure demonstrates that the precision of the evolvent is depend on the density level $m$ used in the construction.
+
+\begin{figure}
+\center
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (b)}
+\end{minipage}
+\caption{Evolvents in two dimensions with (a) $m=4$ and (b) $m=5$}
+\label{fig_evolvent}
+\end{figure}
+
+Отметим, что многие известные алгоритмы глобальной оптимизации неявно основаны на идее редукции размерности и адаптации одномерных алгоритмов для решения многомерных задач \cite{Sergeyev2006,Zilinskas2008,Evtushenko2013}.
+
+%связка
+Правила алгоритма глобального поиска, в котором на каждой итерации проводится параллельно $p$ испытаний, состоят в следующем.
+На первой итерации метода испытания проводятся параллельно в двух граничных точках $x^1 = 0, \; x^2 = 1,$, а также в $(p-2)$ произвольных внутренних точках отрезка $[0,1]$, т.е.
+$x^i\in(0,1),3\leq i \leq p$.
+
+Suppose $n\geq 1$  iterations of the method have been carried out, in which trials were performed at $k=k(n)$ points $x^i,1\leq i \leq k$.
+Then points $x^{k+1},\dots,x^{k+p}$  of search trials of the next iteration are determined according to the following rules.
+
+Rule 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the
+lower indices in ascending order of coordinate values, i.e.
+\[
+0=x_1<x_2<\dots <x_{k-1} <x_k=1,
+\]
+and juxtapose to them the values $z_i=\varphi(y(x_i)), \; 1 \leq i \leq k$, computed at these points.
+
+Rule 2. Compute the current lower estimates
+\begin{equation}\label{Rule_Mu}
+\mu = \max\left\{ \frac{\left|z_i-z_{i-1}\right|}{ \Delta_i },\; 2 \leq i \leq k  \right\},
+\end{equation}
+where $\Delta_i = (x_i-x_{i-1})^{1/N}$. If $\mu$ is equal to zero, then assume $\mu = 1$.
+
+Rule 3. For each interval ($x_{i-1},x_i), \; 2 \leq i \leq k,$ compute
+the \textit{characteristics} $R(i)$ :
+\begin{equation}\label{Rule_R}
+R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r^2 \mu^2\Delta_i}-2\frac{z_i+z_{i-1}}{r \mu},
+\end{equation}
+where $\Delta_i=(x_i-x_{i-1})^{1/N}$. The value $r > 1$ is parameter of the algorithm. An appropriate selection
+of $r$ allows to consider the product $r \mu$ as an estimate
+of the objective function Lipschitz constant $L$.
+
+Rule 4. Arrange characteristics  $R(i), 2 \leq i \leq k$, in decreasing order 
+\begin{equation}\label{Rule_Max}
+R(t_1)\geq R(t_2)\geq \dots \geq R(t_{k}) \geq R(t_{k})
+\end{equation}
+and select $p$ maximum characteristics with interval numbers $t_j, 1\leq j \leq p$.
+
+Rule 5. Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), \; 1\leq j\leq p$, calculated using the formulae
+\begin{equation}\label{Rule_X}
+x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - \frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N.
+\end{equation}
+
+The algorithm terminates if the condition $\Delta_{t_j}<\epsilon$ is satisfied at least for one number $t_j, 1 \leq j \leq p$ ; $\epsilon>0$ is the predefined accuracy.
+
+This method of parallelization has the following justification.
+The characteristics of intervals (\ref{Rule_R}) used in the method can be considered as probability measures of the global minimum point location in these intervals.
+Inequalities (\ref{Rule_Max}) arrange intervals according to their characteristics, and trials are carried out in parallel in the first $p$ intervals with the largest probabilities.
+
+%Параллельный алгоритм - пример работы?
 
@@
 %Локальные задачи и локальные методы
+
+В данном разделе рассмотрим вопросы, связанные с решением подзадачи (\ref{local_problem}), которой (при обозначении $f(x) = f(x,y)$ для фиксированного значения $y\in D$) соответствует задача на локальный экстремум
+
+\begin{equation} \label{lp}
+f(x) \rightarrow \min, x\in S. 
+\end{equation}
+
+К настоящему времени разработано огромное количество разнообразных методов локальной оптимизации для задач вида (\ref{lp}).
+К их числу относятся, например, градиентные методы, ньютоновские и квазиньютоновские методы, методы сопряженных направлений.
+Большинство из них используют принцип локального спуска, когда метод последовательно на каждом шаге переходит к точкам с меньшими значениями целевой функции.
+Почти все эти методы могут быть представлены в виде итерационного соотношения
+\[
+x^{k+1} = x^k + s^k d^k,
+\]
+где $x^k$ -- точки основных испытаний, состоящих в вычислении набора тех или иных локальных характеристик целевой функции $I^k=I(x^k)$ в точке $x^k$, $d^k$ -- направления смещения из точек $x^k$, вычисляемые по результатам основных испытаний, а $s^k$ -- коэффициенты, определяющие величины смещений вдоль выбранных направлений.
+
+Для определения величин смещений $s^k$ вдоль направлений $d^k$ методы могут выполнять вспомогательные (рабочие) шаги.
+Это приводит к дополнительным измерениям локальных характеристик целевой функции вдоль направления $d^k$.
+Переходы от точек $x^k$ к точкам $x^{k+1}$ выполняются таким образом, чтобы обеспечить существенное убывание значений функции $f^k = f( x^k )$ в результате шага.
+
+В набор вычисляемых локальных характеристик $I^k=I(x^k)$ могут входить: значение функции $f^k = f( x^k )$, вектор градиента $\nabla f^k = \nabla f(x^k)$, матрица вторых производных (гессиан) $\Gamma_k=\Gamma^f(x^k)$.
+Какой именно набор характеристик измеряется -- зависит как от свойств решаемой задачи, так и от выбранного метода оптимизации.
+
+Однако в задачах локальной оптимизации, возникающих в приложениях, априорная информация о функциях бывает достаточно ограниченной.
+Например, может предполагаться лишь локальный характер зависимости функции от параметров, тогда как значения градиента (а, тем более, и гессиана) являются неизвестными.
+
+В этом случае применяются методы прямого поиска, которые не используют каких-либо предположений о гладкости минимизируемой функции.
+При поиске минимума эти методы измеряют только значения функции.
+Поскольку информации о гладкости функции, то при выборе направлений смещения методы не могут использовать аппроксимаций функции по результатам ее измерения.
+Правила размещения измерений в них основываются на некоторых эвристических логических схемах.
+
+Одними из популярных методов прямого поиска являются Hooke--Jeeves method \cite{HookJeeves} and Nelder--Mead method \cite{NelderMead}.
+Несмотря на кажущуюся простоту и теоретическую необоснованность методов прямого поиска, они хорошо зарекомендовали себя в реальных расчетах.
+Это можно объяснить следующим образом.
+Многие методы гладкой оптимизации чрезвычайно чувствительны к наличию вычислительных ошибок в значениях функций, превращающих теоретически гладкую функцию в фактически негладкую.
+За счет этого в реальных расчетах они зачастую утрачивают те положительные свойства, которые для них обещает теория.
+Использование методов прямого поиска позволяет в этих условиях добиться лучших результатов.
+
 %Стандартные библиотеки - будем использовать одну из них

### 2019-11-06 2020/PAVT_Nested_Full/pavt_paper.tex
Some corrections in Russian

@@
 Задачи глобальной оптимизации часто возникают в различных областях современной науки и техники.
-Например, в биологии и химии актуальны проблемы поиска конфигураций различных химических соединений, соответствующих минимуму энергии взаимодействия.
+Например,  актуальными являются проблемы поиска конфигураций различных химических соединений, соответствующих минимуму энергии взаимодействия \cite{Posypkin2014}.
 В частности, такие задачи возникают при разработке новых лекарственных препаратов \cite{Sulimov}.
+Задачи глобальной оптимизации возникают и во многих других приложениях (см.
+обзор  \cite{Pinter2006}).
 
@@
 В задачах подобной размерности регулярные методы поиска глобального решения не могут быть применены из-за чрезвычайно больших вычислительных затрат на покрытие области точками испытаний.
-Это остается справедливым даже в случае использования эффективных параллельных алгоритмов, строящих существенно неравномерные покрытия области \cite{Paulavicius2011,Evtushenko2009,Jones2009}.
+Это остается справедливым даже в случае использования эффективных алгоритмов (например, \cite{Paulavicius2011,Evtushenko2009,Jones2009}), строящих существенно неравномерные покрытия области .
 
@@
 Section 3 посвящен описанию параллельного алгоритма глобального поиска с использованием с использованием space-filling curves. Приведена формулировка параллельного алгоритма, пример его работы.
-Section 4 обсуждаются методы решения локальных подзадач в рамках схемы рекурсивной оптимизации. Дается кратное описание свойств задач и известных методов их решения. 
+Section 4 обсуждаются методы решения локальных подзадач в рамках схемы рекурсивной оптимизации. Дается кратное описание свойств задач поиска локального экстремума и известных методов их решения. 
 Section 5 содержит результаты численных экспериментов. Здесь производится оценка масшабируемости предложенного параллельного алгоритма при решении серии тестовых задач.  
@@
 & f(x^\ast,y^\ast)=\min{\left\{f(x,y):x\in S, y\in D\right\}}, \nonumber \\
-& S=\left\{x\in R^M: a_i\leq x_i \leq b_i, 1\leq i \leq M\right\}, \nonumber \\
+& S=\left\{x\in R^M: a_i\leq x_i \leq b_i, 1\leq i \leq M\right\}, \\
 & D=\left\{y\in R^N: c_i\leq y_i \leq d_i, 1\leq i \leq N\right\}. \nonumber
@@
 with the constant $L$ unknown a priori.
-Одновременно с этим при любом фиксированном наборе значений $\overline{y}$  функция $f(x,\overline{y})$ является унимодальной, т.е. функция $f(x,\overline{y})$ имеет единственную точку минимума $x^*$.
+Одновременно с этим при любом фиксированном наборе значений $\overline{y}$  функция $f(x,\overline{y})$ является унимодальной, т.е.
+функция $f(x,\overline{y})$ имеет единственную точку минимума $x^*$ (зависящую, вообще говоря, от $\overline{y}$).
 
 Учет подобной особенности рассматриваемого класса задач может существенно снизить вычислительную сложность поиска оптимума.
-В самом деле, в соответствии с известной схемой рекурсивной оптимизации (ссылка) решение исходной задачи можно свести к задаче поиска глобального минимума функции $\varphi(y)$
+В самом деле, в соответствии с известной схемой рекурсивной оптимизации \cite{Carr} решение исходной задачи можно свести к задаче поиска глобального минимума функции $\varphi(y)$
 \begin{equation}\label{global_problem}
@@
 \end{equation}
-В соответствии с данной схемой вычисление одного значения функции $\varphi (y)$ (данный процесс для задачи \ref{global_problem} будем называть search trial) подразумевает решение унимодальной задачи (\ref{local_problem}), которое может быть выполнено одним из методов локальной оптимизации.
-Эффективность и проработанность методов локальной оптимизации позволяет решать подзадачи (\ref{local_problem}) с высокой точностью.
-Соответственно, можно рассматривать задачу (\ref{global_problem}) как задачу глобальной оптимизации, в которой значение функции может быть вычислено точно, но данная операция является трудоемкой.
+В соответствии с данной схемой вычисление одного значения функции $\varphi (y)$ (данный процесс будем называть search trial) подразумевает решение унимодальной задачи (\ref{local_problem}), которое может быть выполнено одним из методов локальной оптимизации.
+Эффективность и проработанность методов локальной оптимизации позволяет решать подзадачи (\ref{local_problem}) с точностью, значительно превышающей точность методов глобальной оптимизации.
+Соответственно, можно рассматривать задачу (\ref{global_problem}) как задачу глобальной оптимизации, в которой значение функции может быть вычислено с высокой точностью, но данная операция является трудоемкой.
 
@@
 \[
-\Delta(\omega,c)= \sum^{m}_{j=1}\left[u_j-u(\tau_j,\omega,c)\right]^2. 
+\Delta(\omega,c)= \sum^{m}_{j=1}\left[u_j-u(t_j,\omega,c)\right]^2. 
 \]
@@
 Вложенная подзадача (\ref{inner_prob}) является классической линейной задачей наименьших квадратов, и ее решение can be obtained by solving a system of linear algebraic equations regarding the unknown $c$, which can be done, e.g., by Gaussian elimination.
-Одновременно с этим внешняя задача (\ref{outer_prob}) будет являться, вообще говоря, многоэкстремальной.
+Одновременно с этим внешняя задача (\ref{outer_prob}) будет являться, вообще говоря, многоэкстремальной и ее решение потребует значительных вычислительных ресурсов.
 
@@
 Лобачевского был предложен \cite{Barkalov2018,Strongin2018,globalizerSystem} эффективный параллельный алгоритм глобальной оптимизации, предназначенный для решения задач вида (\ref{global_problem}) .
-Основная идея в распараллеливании состоит в такой организации вычислений, при которой параллельно проводятся несколько испытаний.
+Основная идея распараллеливания состоит в такой организации вычислений, при которой параллельно проводятся несколько испытаний.
 Этот подход характеризуется эффективностью (распараллеливается именно та часть вычислительного процесса, в котором выполняется основной объем вычислений) и общностью (применим для широкого класса алгоритмов глобальной оптимизации).
@@
 Правила алгоритма глобального поиска, в котором на каждой итерации проводится параллельно $p$ испытаний, состоят в следующем.
-На первой итерации метода испытания проводятся параллельно в двух граничных точках $x^1 = 0, \; x^2 = 1,$, а также в $(p-2)$ произвольных внутренних точках отрезка $[0,1]$, т.е.
+На первой итерации метода испытания проводятся параллельно в двух граничных точках $x^1 = 0, \; x^2 = 1$, а также в $(p-2)$ произвольных внутренних точках отрезка $[0,1]$, т.е.
 $x^i\in(0,1),3\leq i \leq p$.
@@
 
-%Параллельный алгоритм - пример работы?
+Various modifications of this algorithm and the corresponding theory of
+convergence are given in \cite{Strongin2000,Barkalov2002,Gergel2005,Strongin2013}.
+
+%Параллельный алгоритм - пример работы? Нужен ли?
 
@@
 
-В данном разделе рассмотрим вопросы, связанные с решением подзадачи (\ref{local_problem}), которой (при обозначении $f(x) = f(x,y)$ для фиксированного значения $y\in D$) соответствует задача на локальный экстремум
+В данном разделе рассмотрим вопросы, связанные с решением подзадачи (\ref{local_problem}), которой (при обозначении $f(x) = f(x,\overline{y})$ для фиксированного значения $\overline{y} \in D$) соответствует задача на локальный экстремум
 
@@
 
-Однако в задачах локальной оптимизации, возникающих в приложениях, априорная информация о функциях бывает достаточно ограниченной.
+В задачах локальной оптимизации, возникающих в приложениях, априорная информация о функциях бывает достаточно ограниченной.
 Например, может предполагаться лишь локальный характер зависимости функции от параметров, тогда как значения градиента (а, тем более, и гессиана) являются неизвестными.
-
 В этом случае применяются методы прямого поиска, которые не используют каких-либо предположений о гладкости минимизируемой функции.
 При поиске минимума эти методы измеряют только значения функции.
-Поскольку информации о гладкости функции, то при выборе направлений смещения методы не могут использовать аппроксимаций функции по результатам ее измерения.
-Правила размещения измерений в них основываются на некоторых эвристических логических схемах.
+Правила размещения точек итераций в них основываются на некоторых эвристических логических схемах.
 

### 2019-11-06 2019/NUMTA_Barkalov/Special Issure/article.tex
Template for special issure

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Journal of \LaTeX\ Templates}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{Elsevier\fnref{myfootnote}}
\address{Radarweg 29, Amsterdam}
\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
\ead[url]{www.elsevier.com}

\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{support@elsevier.com}

\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

\begin{keyword}
\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{The Elsevier article class}

\paragraph{Installation} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.

\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript. Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.

\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class. In addition, it features commands and options to format the
\begin{itemize}
\item document style
\item baselineskip
\item front matter
\item keywords and MSC codes
\item theorems, definitions and proofs
\item lables of enumerations
\item citation style and labeling.
\end{itemize}

\section{Front matter}

The author names and affiliations could be formatted in two ways:
\begin{enumerate}[(1)]
\item Group the authors per affiliation.
\item Use footnotes to indicate the affiliations.
\end{enumerate}
See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.

\section{Bibliography styles}

There are various bibliography styles available. You can select the style of your choice in the preamble of this document. These styles are Elsevier styles based on standard styles like Harvard and Vancouver. Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.

Here are two sample references: \cite{Feynman1963118,Dirac1953888}.

\section*{References}

\bibliography{bibliography}

\end{document}

### 2019-11-13 2017/AMS_IsrafilovGrishagin/paper.tex
Paper from AMC2017 by Grishagin and Israfilov

\documentclass[preprint]{elsarticle}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{csquotes}
\usepackage{longtable}
\usepackage{microtype}

\usepackage{tikz}
\usetikzlibrary{trees, positioning}
\tikzset{
  edge from parent/.style = {draw,
    edge from parent fork right
  },
  basic/.style = {draw,
    align = center,
    font=\fontsize{8}{8}\selectfont
  },
  dots/.style = {
    font=\large
  },
  dots_large/.style = {
    font=\LARGE
  },
  root/.style = {basic},
  level 1/.style = {basic,
    text width = 75pt,
    level distance = 95pt,
    sibling distance = 80pt
  },
  level 2/.style = {basic,
    text width = 110pt,
    level distance = 130pt,
    sibling distance = 42pt
  }
}

\newtheorem{theorem}{Theorem}
\newproof{proof}{Proof}

\DeclareMathOperator{\sign}{sign}

\biboptions{numbers,sort&compress}
\bibliographystyle{elsarticle-num}

\begin{document}

\begin{frontmatter}

\title{Convergence conditions and numerical comparison of global optimization methods based on dimensionality reduction schemes}

% Authors
\author[russia_address]{Vladimir Grishagin\corref{cor_vagris}}
\ead{vagris@unn.ru}

\author[russia_address]{Ruslan Israfilov}
\ead{ruslan@israfilov.com}

\author[russia_address,italy_address]{Yaroslav Sergeyev}
\ead{yaro@dimes.unical.it}

\address[russia_address]{Department of Software and Supercomputing, Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia.}
\address[italy_address]{DIMES, Università della Calabria, Italy.}

\cortext[cor_vagris]{Corresponding author}

\begin{abstract}
This paper is devoted to numerical global optimization algorithms applying several ideas to reduce the problem dimension. Two approaches to the dimensionality reduction are considered. The first one is based on the nested optimization scheme that reduces the multidimensional problem to a family of one-dimensional subproblems connected in a recursive way. The second approach as a reduction scheme uses Peano-type space-filling curves mapping multidimensional domains onto one-dimensional intervals. In the frameworks of both the approaches, several univariate algorithms belonging to the characteristical class of optimization techniques are used for carrying out the one-dimensional optimization. Theoretical part of the paper contains a substantiation of global convergence for the considered methods. The efficiency of the compared global search methods is evaluated experimentally on the well-known GKLS test class generator used broadly for testing global optimization algorithms. Results for representative problem sets of different dimensions demonstrate a convincing advantage of the adaptive nested optimization scheme with respect to other tested methods.
\end{abstract}

\begin{keyword}
Multiextremal functions, global optimization, numerical methods, dimensionality reduction, convergence, comparison of efficiency.
\MSC[2010] 65K05\sep 90C26
\end{keyword}

\end{frontmatter}

\section{Introduction}
In this paper the black-box global optimization problem
\begin{gather}
  \label{eq:go_problem}
  f^* = f(y^*) = \min_{y \in P} f(y), \\
  \label{eq:hyper}
  P = \{ y \in \mathbb{R}^N \colon a_i \leq y_i \leq b_i, \; 1 \leq i \leq N \},
\end{gather}
%
is considered as a problem of finding the global minimum value $f^*$ and global minimizers $y^* \in P$ of a real-valued multivariate function $f(y)$ in the hyperparallelepiped~(\ref{eq:hyper}) of the Euclidean space $\mathbb{R}^N$. The objective function $f(y)$ is supposed to satisfy in the domain $P$ the Lipschitz condition
\begin{equation}
  \label{eq:lipschitz}
  |f(y') - f(y'')| \leq L\| y' - y'' \|, \; y', y'' \in P,
\end{equation}
%
where $L > 0$ is a finite constant and $\|*\|$ denotes the Euclidean norm in $\mathbb{R}^N$. In general case, these problems are multiextremal and non-smooth.

The global optimization problem under consideration has been drawing attention of many researchers (see, for example, fundamental monographs~\cite{bib1,bib2,bib3,bib4,bib5,bib6}). On the one hand, problems of this kind are very important from the practical point of view because they arise often in scientific and engineering applications (see, for instance,~\cite{bib4,bib7,bib8,bib9,bib10,bib11,bib12,bib13}). On the other hand, Lipschitzian problems~(\ref{eq:go_problem})--(\ref{eq:hyper}) are very interesting for theoretical study because they have a rich variety of properties and, as a consequence, there is no a \enquote{universal} algorithm for solving multiextremal problems. These circumstances generate many fruitful approaches for solving this class of problems. Given approaches are based on ideas of different nature (both stochastic~\cite{bib5,bib6,bib14,bib16,bib17,bib18} and deterministic~\cite{bib19,bib20,bib21,bib22,bib23,bib24,bib25,bib26,bib27,bib28,bib29,bib30,bib46}), but, in any case, numerical methods of searching for the global optimum are proposed within the frameworks of approaches as a tool for getting a solution sought.

As it was discussed in~\cite{bib5}, the global optimum is an integral characteristic of the problem, i.e., in order to make sure that a point $y^* \in P$ is the global minimizer of the problem~(\ref{eq:go_problem})--(\ref{eq:hyper}) it is required to compare the value $f(y^*)$ with values of the objective function at all points of the domain $P$, but not in a vicinity of $y^*$ only. As a result, when minimizing an essentially multiextremal function, a numerical global optimization method has to build a grid (random or regular) in the feasible domain and the number of grid nodes increases exponentially when rising the problem dimension. This peculiarity causes the substantial complexity of multiextremal problems and dimension is a crucial factor influencing significantly the efficiency of global optimization algorithms.

In this situation, approaches to elaboration of computational schemes reducing the dimension are widely used in the multiextremal optimization. Here we consider two approaches in which the initial multidimensional problem~(\ref{eq:go_problem})--(\ref{eq:hyper}) is reduced to one or several univariate subproblems solved by efficient one-dimensional algorithms. The first approach applies \emph{the nested optimization scheme} in its classical version \cite{bib31,bib32,bib33,bib34,bib35,bib36,bib37} and in its generalization --- \emph{the adaptive nested scheme}~\cite{bib38,bib39}. The second approach is based on \emph{Peano space-filling curves} mapping the multidimensional domain~(\ref{eq:hyper}) onto an interval in one-dimensional space $\mathbb{R}^1$~\cite{bib5,bib40,bib41,bib42,bib43,bib44,bib45,bib47}.

The rest of the paper is organized in the following way. Section~\ref{sec:dim_reduction} describes the basic structures of the reduction schemes mentioned above and the univariate characteristical methods of global optimization to be applied within the reduction structures for solving the internal one-dimensional subproblems. Section~\ref{sec:convergence} is devoted to a theoretical substantiation of the convergence for multidimensional methods combining the reduction schemes with characteristical methods of global search and contains both the known and new theoretical results. Section~\ref{sec:num_exp} presents experimental results of efficiency comparison for the methods described in previous sections on representative sets of multiextremal test functions of various dimensions belonging to the popular test class GKLS~\cite{bib48} with a controllable complexity. Finally, Section~\ref{sec:conclusion} contains a brief conclusion.

\section{Schemes of dimensionality reduction}
\label{sec:dim_reduction}

The first approach to dimensionality reduction called the scheme of nested optimization is based on the well-known relation (see, e.g.,~\cite{bib5,bib26,bib31})
\begin{equation}
  \label{eq:nested_eq}
  \min_{y \in P} f(y) = \min_{y_1 \in [a_1, b_1]} \dots \min_{y_N \in [a_N, b_N]} f(y_1, \dots, y_N).
\end{equation}
%
In order to describe the scheme let us define a family of reduced functions as follows:
\begin{gather}
  f^N(y) \equiv f(y), \\
  f^i(y_1, \dots, y_i) = \min_{y_{i + 1} \in [a_{i + 1}, b_{i + 1}]} f^{i + 1} (y_1, \dots, y_i, y_{i + 1}), \; 1 \leq i \leq N - 1.
\end{gather}
%
Then, according to~(\ref{eq:nested_eq}), in order to find the solution to the multidimensional problem~(\ref{eq:go_problem})--(\ref{eq:hyper}) it is sufficient to solve the one-dimensional problem
\begin{equation}
  \label{eq:subproblem_0}
  f(y^*) = \min_{y_1 \in [a_1, b_1]} f^1(y_1).
\end{equation}
%
But in order to evaluate the function $f^1$ at a fixed point $y_1$ it is necessary to solve the one-dimensional problem of the second level
\begin{equation}
  f^1(y_1) = \min_{y_2 \in [a_2, b_2]} f^2(y_1, y_2),
\end{equation}
%
and so on up to the univariate minimization at the $N$-th level the function $f^N(y) = f(y)$ with fixed coordinates $y_1, \dots, y_{N - 1}$.

Thus, instead of the multidimensional problem~(\ref{eq:go_problem})--(\ref{eq:hyper}) one can solve the set of nested one-dimensional subproblems
\begin{equation}
  \label{eq:subproblem_i}
  \min_{y_i \in [a_i, b_i]} f^i(y_1, \dots, y_{i - 1}, y_i), \; 1 \leq i \leq N,
\end{equation}
%
in which the coordinates $y_1, \dots, y_{i - 1}$ have been already determined by the subproblems of preceding levels. The objective functions $f^i(y_1, \dots, y_i)$ of the subproblems~(\ref{eq:subproblem_i}) satisfy the Lipschitz condition for the corresponding argument $y_i$ subject to the assumption~(\ref{eq:lipschitz}) for the function $f(y)$ (see~\cite{bib51}). Therefore, efficient univariate optimization algorithms can be taken for solving these subproblems. Various combinations of such the algorithms with the nested scheme~(\ref{eq:nested_eq}) have been proposed in~\cite{bib5,bib9,bib26,bib33,bib34,bib35}.

The recursive character of generating the subproblems~(\ref{eq:subproblem_i}) during the optimization forms a hierarchical structure of subordination for these subproblems as a tree where the root is the subproblem~(\ref{eq:subproblem_0}) and the leaves are the
subproblems of the $N$-th level. Fig.~\ref{fig:nested_scheme} shows such a structure for $N = 3$.

Algorithmic implementation of the nested optimization scheme has some features. In the classical version (see, e.g.,~\cite{bib5,bib31}) during solving the root problem~(\ref{eq:subproblem_0}) a new evaluation of the function $f^1$ at a new point (new search iteration), i.e., solving a new subproblem of minimization of $f^2$ at the second level is supposed to be carried out only after the completion of the current iteration. It means that only one minimization subproblem at the second level can be active (can be in the course of solving) because the other subproblems of this level either have been already completed or will be initiated later. Analogously, the same situation takes place at all the rest of levels and at every moment of the multidimensional optimization not more than $N$ univariate subproblems are active and they belong to a path from the root to a leaf. In Fig.~\ref{fig:nested_scheme} one of the possible paths consists of subproblems marked with grey colour. Such organization of computations signifies that the information on the subproblems which have been already completed at the level $i$ is not used during current minimization of a function $f^i$ from~(\ref{eq:subproblem_i}). For example, in Fig.~\ref{fig:nested_scheme} all the subproblems located above the active ones (marked with grey) have been already solved and information on the objective function obtained in the course of solving these subproblems is not used in the currently solved subtasks. Such loss of information leads to increasing the number of trials (evaluations of the objective functions) and worsens the efficiency of optimization. Indeed, if during the minimization of a function $f^i$ from~(\ref{eq:subproblem_i}) we get values higher than ones obtained in a subproblem of the same level solved earlier, then, may be, it is reasonable not to continue solving the current subproblem up to the end and not to spend excessive trials.

This idea is realized \emph{in the adaptive nested optimization scheme} proposed in the paper~\cite{bib38} containing a detailed algorithmic description of the adaptive scheme. The main suggestion is to give up the principle of strict subordination inherent to the classical nested scheme and to use a \emph{simultaneous consideration} of all the univariate subproblems arising in the course of multidimensional optimization. In the adaptive scheme all the generated subproblems are active and a numerical \enquote{measure of quality} for each subproblem is introduced (this measure is called the \emph{subproblem characteristic}). An iteration of the multidimensional optimization consists in the choice of the subproblem with maximal characteristic and carrying out a new trial within this subproblem. Such the approach allows choosing the subproblems with lower values of objective univariate functions~(\ref{eq:subproblem_i}) more often than with the higher ones. The way of defining the subproblem characteristics can be different, but in the case when the univariate characteristical algorithms~\cite{bib49,bib50} are used for solving the subproblems~(\ref{eq:subproblem_i}) (see the general description of these algorithms below) the maximal interval characteristic generating by the algorithm in the course of optimization can be taken as the subproblem characteristic.

The second approach is based on the known fact that a finite interval $[a, b]$ of the real axis and the $N$-dimensional hyperparallelepiped~(\ref{eq:hyper}) are equipollent sets and there exist mappings called \emph{Peano space-filling curves} (or \emph{evolvents}) which map the unit interval $[a, b]$ onto the hyperparallelepiped~(\ref{eq:hyper}) continuously and unambiguously.

Let $y(x)$ be such an evolvent. Let us consider the function
\begin{equation}
  \label{eq:peano_map}
  \varphi(x) = f(y(x))
\end{equation}
%
and the one-dimensional problem
\begin{equation}
  \label{eq:peano_min}
  \varphi^* = \varphi(x^*) = \min_{x \in [a, b]} \varphi(x).
\end{equation}
%
Owing to~(\ref{eq:lipschitz}) the function $f(y)$ is continuous and, because of continuity of $f(y)$ and $y(x)$,
\begin{equation}
  \label{eq:peano_min_2}
  \min_{y \in P} f(y) = \min_{x \in [a, b]} \varphi(x),
\end{equation}
%
and the point $y^* = y(x^*)$ is the global minimizer of the multidimensional problem~(\ref{eq:go_problem})--(\ref{eq:hyper}).

As a result, solving the univariate problem~(\ref{eq:peano_min}) allows obtaining the solution to the initial problem~(\ref{eq:go_problem})--(\ref{eq:hyper}). A wide spectrum of the methods exploiting this idea is presented, for example, in publications~\cite{bib5,bib29,bib40,bib43,bib44,bib45,bib47}. It should be noticed that the objective function $\varphi(x) = f(y(x))$ of the univariate problem~(\ref{eq:peano_min}) satisfies in general case the H\"older condition
\begin{equation}
  \label{eq:holder}
  | \varphi(x') - \varphi(x'') | \leq H \sqrt[N]{|x' - x''|}, \; x', x'' \in [a, b],
\end{equation}
%
if the function $f(y)$ is Lipschitzian~(\ref{eq:lipschitz}), and the resulting H\"older constant $H$ is finite and depends on the Lipschitz
constant $L$ and the dimension $N$~\cite{bib5,bib51} only.

Combinations of the reduction schemes described above with different univariate techniques of global search generate corresponding multidimensional algorithms of solving the multiextremal problem~(\ref{eq:go_problem})--(\ref{eq:hyper}). Since in the nested optimization schemes (both classical and adaptive) all the univariate subproblems~(\ref{eq:subproblem_i}) satisfy the Lipschitz condition, we considered as one-dimensional methods the known algorithms of Lipschitzian optimization proposed by Piyavskij~\cite{bib26} and Strongin~\cite{bib15} which are theoretically substantiated and quite efficient. In regard to the reduction on the base Peano evolvents, the reduced objective function~(\ref{eq:peano_map}) is H\"olderian (\ref{eq:holder}) and for solving the univariate problem (\ref{eq:peano_min_2}) it is necessary to apply the methods oriented at this class of optimization problems. As such the methods we took two algorithms which are generalizations~\cite{bib5,bib29} of the Piyavskij's and Strongin's methods mentioned above.

All the univariate methods chosen for our research belong to the class of characteristical algorithms introduced by Grishagin~\cite{bib50} (see also~\cite{bib49}). Therefore, at first we describe a general structure of the characteristical methods and then give in details the specificity of each method in the framework of the general description.

Let us consider the problem of univariate optimization in a standardized form: to find the minimal value
\begin{equation}
  \label{eq:univariate}
  \varphi^* = \varphi(x^*) = \min_{x \in [a, b]} \varphi(x)
\end{equation}
%
of a function $\varphi(x)$ over a finite interval $[a, b] \subset \mathbb{R}^1$ where the objective function $\varphi(x)$ satisfies either Lipschitz~(\ref{eq:lipschitz}) or H\"older~(\ref{eq:holder}) condition in the feasible domain $[a,b]$.

Let the term \enquote{trial} denote an operation that calculates the value of the function $\varphi(x)$ at a point $x \in [a, b]$. An optimization method is supposed to place sequentially trials at points $x^1, x^2, \dots, x^k, \dots $ with corresponding outcomes $z^k = \varphi(x^k)$, $k = 1, 2, \dots$.

A numerical method of solving the problem~(\ref{eq:univariate}) is characteristical if its decision rule of planning trials consists of the following actions.

The first two trials are to be carried out at the end-points of the interval $[a, b]$, i.e.,
\begin{equation}
  x^1 = a, \; x^2 = b,
\end{equation}
with the values $z^1 = \varphi(a)$, $z^2 = \varphi(b)$.

The point $x^{k + 1}$, $k > 2$, of any subsequent $(k + 1)$-th trial is chosen in accordance with the following steps:
\begin{enumerate}[Step 1.]
  % Change spacing between items due to text blending
  \setlength\itemsep{0.5em}

  \item All the points $x^1, x^2, \dots, x^k$ of preceding trials are renumbered by subscripts in increasing order, i.e.,
  \begin{equation}
    \label{eq:x_sorted}
    a = x_1 < x_2 < \dots < x_k = b.
  \end{equation}
  %
  The values $z_i = \varphi(x_i)$, $1 \leq i \leq k$, are juxtaposed to the points $x_i$, $1 \leq i \leq k$, from~(\ref{eq:x_sorted}).

  \item For each subinterval $(x_{i - 1}, x_i)$, $2 \leq i \leq k$, a numerical value $R(i)$ is calculated (this value is called characteristic of the subinterval).

  \item The subinterval $(x_{t - 1}, x_t)$ with the maximal characteristic is selected among all the subintervals:
  \begin{equation}
    \label{eq:r_max}
    R(t) = \max_{2 \leq i \leq k} R(i).
  \end{equation}

  \item The next $(k + 1)$-th trial is executed at a point $x^{k + 1} \in (x_{t - 1}, x_t)$ and the value $z^{k + 1} = \varphi(x^{k + 1})$ is calculated.
\end{enumerate}

The general conditions of convergence to global minimum for characteristical algorithms have been proven in~\cite{bib49}. In particular, these results substantiate the termination criterion in the form
\begin{equation}
  \label{eq:stop_criterion}
  x_t - x_{t - 1} \leq \varepsilon,
\end{equation}
%
where $t$ is the argument of the largest characteristic $R(t)$ from~(\ref{eq:r_max}) and $\varepsilon > 0$ is a predefined coordinate accuracy of the search, i.e., rules 1--4 are carried out until the length of the interval with maximal characteristic is
less than the accuracy $\varepsilon$.

Now in order to describe a characteristical algorithm it is sufficient to define the expressions for its characteristics and the rule of placement of the next trial within the interval with maximal characteristic.

Following this way, we give corresponding descriptions for 4 univariate characteristical methods used in the reduction schemes under consideration.

The first two methods are intended for optimization of univariate Lipschitzian functions and are used in classical and adaptive nested schemes.

The first method was proposed by Piyavskij~\cite{bib26} and it is well-known in the Lipschitzian optimization. Its characteristics $R(i)$ are calculated as
\begin{equation}
  R(i) = \frac{m (x_i - x_{i - 1})}{2} - \frac{z_{i - i} + z_i}{2},
\end{equation}
%
and the point of new trial is calculated in accordance with the expression
\begin{equation}
  \label{eq:new_trial}
  x^{k + 1} = \frac{x_{t - 1} + x_t}{2} - \frac{z_t - z_{t - 1}}{2m},
\end{equation}
%
where $m > 0$ is the parameter of the method. We will use for short designation of this algorithm the abbreviation
PM (\emph{Piyavskij's Method}).

The second algorithm was proposed by Strongin~\cite{bib5,bib15,bib51}. It has the characteristics
\begin{equation}
  R(i) = m (x_i - x_{i - 1}) + \frac{(z_i - z_{i - 1})^2}{m (x_i - x_{i - 1})} - 2 (z_{i - 1} + z_i),
\end{equation}
%
and its new trial point $x^{k + 1}$ is given in accordance with the relation~(\ref{eq:new_trial}). The value $m > 0$ is the parameter of the method. The author of the method called it simply \emph{Global Search Algorithm} and we keep this name in the short form GSA.

If the objective function in the problem~(\ref{eq:univariate}) satisfies the Lipschitz condition with the Lipchitz constant $L > 0$ then the convergence to all global minima is provided by the methods described when the parameter $m > L$ for PM and $m > 2L$ in GSA. Unfortunately, the value of the Lipchitz constant is unknown, as a rule, and for the parameter $m$ one can use its adaptive estimation~\cite{bib5,bib51}.
\begin{equation}
  \label{eq:m}
  m =
  \begin{cases}
    rM, & M > 0, \\
    1,  & M = 0,
  \end{cases}
\end{equation}
%
where
\begin{equation}
  \label{eq:M}
  M = \max_{2 \leq i \leq k} \frac{|z_i - z_{i - 1}|}{x_i - x_{i - 1}}
\end{equation}
%
and $r > 1$ is the parameter of the method.

The next two algorithms are oriented at solving the problem~(\ref{eq:peano_min_2}) with a H\"olderian objective function and are generalizations of PM and GSA to this case. The third methods called GPM (Generalized Piyavskij's Method) uses the characteristics
\begin{equation}
  R(i) = \frac{m \sqrt[N]{x_i - x_{i - 1}}}{2} - \frac{z_{i - 1} + z_i}{2},
\end{equation}
%
and the new trial point
\begin{equation}
  \label{eq:general_new_trial}
  x^{k + 1} = \frac{x_{t - 1} + x_t}{2} - \sign(z_t - z_{t - 1}) \frac{|z_t - z_{t - 1}|^N}{2 r m^N}.
\end{equation}

The fourth method --- Generalized Global Search Algorithm (GGSA) contains the characteristics
\begin{equation}
  R(i) = m \sqrt[N]{x_i - x_{i - 1}} + \frac{(z_i - z_{i - 1})^2}{m \sqrt[N]{x_i - x_{i - 1}}} - 2 (z_{i - 1} + z_i),
\end{equation}
%
and the point of the new trial is calculated in accordance with~(\ref{eq:general_new_trial}).

In both the methods the parameter $m$ is taken in the form~(\ref{eq:m}), but instead of~(\ref{eq:M}) the relation
\begin{equation}
  M = \max_{2 \leq i \leq k} \frac{|z_i - z_{i - 1}|}{\sqrt[N]{x_i - x_{i - 1}}},
\end{equation}
%
is used.

\section{Convergence of multidimensional reduction algorithms}
\label{sec:convergence}
Combinations of the nested dimensionality reduction schemes with the univariate methods PM and GSA generate the multidimensional algorithms for solving the problem~(\ref{eq:go_problem}). Let us call these algorithms in the following manner: PM-C and GSA-C are the classical nested scheme with the methods PM and GSA correspondingly at the univariate levels of reduction~(\ref{eq:subproblem_i}), and PM-A and GSA-A realize the adaptive nested scheme in which for the univariate optimizations~(\ref{eq:subproblem_i}) the methods PM and GSA are used respectively.

Convergence to the global solution of the problem~(\ref{eq:go_problem}) was proven for GSA-C in the monograph~\cite{bib51} and for GSA-A in the paper~\cite{bib38}. Moreover, the methods GPM and GGSA provide the property of global convergence as well (see~\cite{bib29,bib51}). As for PM-C and PM-A, the problem of their convergence to absolute minimum was open so far, and in the present paper we give a theoretical substantiation of this important property.

First of all, we prove an auxiliary assertion on the sensitivity of PM to the possible errors in calculations of objective function values.

Describing the general structure of the characteristical algorithms we supposed that the results $z^k$ at the trial points coincide with strict values of the objective function~(\ref{eq:univariate}), i.e., $z^k = \varphi(x^k)$, $k = 0, 1, \dots$. But in the real optimization there are disturbances caused by computational errors. Assume that these errors are bounded by a real number $\xi > 0$. Then~\cite{bib51} the results $z^k$ of trials can be treated as precise values of a function $\omega(x)$ such that
\begin{equation}
  \label{eq:convergence_cond}
  | \omega(x) - \varphi(x) | \leq \xi, \; x \in [a, b].
\end{equation}

After termination of the search caused by the stopping criterion~(\ref{eq:stop_criterion}) the value
\begin{equation}
  \omega_k^* = \min_{1 \leq i \leq k} z^i
\end{equation}
%
may be considered as an approximation of the global minimum $\varphi^*$ of the function $\varphi(x)$ from~(\ref{eq:univariate}). In this case,
the sensitivity of the method to the bounded errors can be estimated by a degree of proximity between $\omega_k^*$ and $\varphi^*$. The next theorem provides such an estimation for the algorithm PM.

\begin{theorem}
  \label{thm:theorem1}
  Assume that after a termination of the search according to the stopping criterion~(\ref{eq:stop_criterion}) there is a truncated sequence of trials $x^1, x^2 \dots, x^k$ generated by PM during minimizing the approximation $\omega(x)$ of the objective function $\varphi(x),~x\in [a,b]$, satisfying the inequality~(\ref{eq:convergence_cond}), and the function $\varphi(x)$ is Lipschitzian with constant $L > 0$. Then the following inequality holds:
  %
  \begin{equation}
    \label{eq:theorem_1_cond}
    | \omega_k^*  - \varphi^* | \leq \frac{L \varepsilon}{2} + \xi,
  \end{equation}
  if the PM's parameter $m > L$.
\end{theorem}

\begin{proof}
  Let us order the elements of the truncated sequence $x^1, x^2, \dots, x^k$ in accordance with the rule~(\ref{eq:x_sorted}) and introduce the designations $\alpha_i = (z_{i - 1} + z_i) / 2$, $\delta_i = x_i - x_{i - 1}$, $2 \leq i \leq k$, where $z_i = \omega(x_i)$, $1 \leq i \leq k$.

  From the Lipschitz condition for the function $\varphi(x)$
  \begin{equation}
    \varphi(x_i) + \varphi(x_{i - 1}) - L\delta_i \leq 2\varphi^*, \; 2 \leq i \leq k,
  \end{equation}
  %
  and owing to~(\ref{eq:convergence_cond})
  \begin{equation}
    z_i - \varphi(x_i) \leq \xi, \; z_{i - 1} - \varphi(x_{i - 1}) \leq \xi, \; 2 \leq i \leq k,
  \end{equation}
  %
  from where
  \begin{equation}
    \alpha_i - \frac{L \delta_i}{2} \leq \xi + \varphi^*, \; 2 \leq i \leq k,
  \end{equation}
  %
  or
  \begin{equation}
    \label{eq:min_alpha_L_delta}
    \min_{2 \leq i \leq k} \Big( \alpha_i - \frac{L \delta_i}{2} \Big) - \xi \leq \varphi^*.
  \end{equation}
  %
  On the other hand, for the point $x_s$ such that $\omega(x_s) = \omega_k^*$ taking into account~(\ref{eq:convergence_cond}) and the inequality $\varphi(x_s) \geq \varphi^*$ we have
  \begin{equation}
    \label{eq:phi_leq_omega_xi}
    \varphi^* \leq \omega_k^* + \xi.
  \end{equation}
  %
  Let the number $q$, $2 \leq q \leq k$, be such that
  \begin{equation}
    \alpha_q - \frac{L \delta_q}{2} = \min_{2 \leq i \leq k} \Big( \alpha_i - \frac{L \delta_i}{2} \Big).
  \end{equation}
  %
  Let us consider two cases.
  The first one corresponds to the situation $\delta_q \leq \varepsilon$ for $\varepsilon$ from~(\ref{eq:stop_criterion}). Since $\omega_k^* \leq \alpha_q$ then
  \begin{equation}
    \omega_k^* - \frac{L \varepsilon}{2} \leq \alpha_q - \frac{L \delta_q}{2} \leq \alpha_i - \frac{L \delta_i}{2}
  \end{equation}
  %
  for all $i$, $2 \leq i \leq k$.

  In the opposite case $\delta_q > \varepsilon$ let us take the interval $(x_t, x_{t - 1})$ with the largest characteristic $R(t)$ from~(\ref{eq:r_max}) which the inequality~(\ref{eq:stop_criterion}) is true for. The characteristic $R(q) \leq R(t)$ because of~(\ref{eq:r_max}), i.e.,
  \begin{equation}
    R(t) - R(q) = \frac{m(\delta_t - \delta_q)}{2} - \alpha_t + \alpha_q \geq 0,
  \end{equation}
  %
  from where
  \begin{equation}
    \alpha_t - \frac{L \delta_t}{2} \leq \alpha_q - \frac{L \delta_q}{2} - \frac{(m - L)(\delta_q - \delta_t)}{2} \leq \alpha_q -\frac{L \delta_q}{2},
  \end{equation}
  %
  taking into account $m > L$ and $\delta_q > \varepsilon \geq \delta_t$.

  But $\alpha_t - L \delta_t / 2 \geq \omega_k^* - L \varepsilon / 2$, therefore, from~(\ref{eq:min_alpha_L_delta}) it follows that
  \begin{equation}
    \label{eq:omega_phi_leq_frac_xi}
    \omega_k^* - \varphi^* \leq \frac{L \varepsilon}{2} + \xi.
  \end{equation}
  %
  On the other hand, due to~(\ref{eq:phi_leq_omega_xi}) we have
  \begin{equation}
    \label{eq:xi_two_sides}
    \varphi^* - \omega_k^* \leq \xi \leq \xi + \frac{L \varepsilon}{2}.
  \end{equation}
  %
  The inequalities~(\ref{eq:omega_phi_leq_frac_xi}),~(\ref{eq:xi_two_sides}) prove the main assertion~(\ref{eq:theorem_1_cond}) of the theorem. The proof has been completed.
\end{proof}

Now let us return to the multidimensional problem~(\ref{eq:go_problem}) and the algorithm PM-C. It should be noted that the univariate subproblems~(\ref{eq:subproblem_i}) are solved approximately. Indeed, the evaluation of the function $f^{N - 1}(y_1,$ $\dots, y_{N - 1})$ for fixed $y_1, \dots, y_{N - 1}$ consists in minimization by means of PM the function $f^N(y_1,$ $\dots, y_{N - 1}, y_N)$ for the coordinate $y_N$ and instead of the precise value $f^{N - 1}(y_1,$ $\dots, y_{N - 1})$ we obtain an approximate estimation $\omega^{N - 1}(y_1, \dots, y_{N - 1})$. Furthermore, evaluation $f^{N - 2}(y_1,$ $\dots, y_{N - 2})$ for fixed $y_1, \dots, y_{N - 2}$ is implemented as an approximate minimization of the function $\omega^{N - 1}(y_1,$ $\dots, y_{N - 1})$ and the result of this minimization is an approximate estimation $\omega^{N - 2}(y_1,$ $\dots, y_{N - 2})$, etc., up to obtaining an estimation $\omega^* = \omega^1$ in the subproblem $\min \{ \omega^1(y_1) : y_1 \in [a_1, b_1] \}$. In this procedure the resulting value $\omega^*$ is an estimation of the sought solution $f^*$ from~(\ref{eq:go_problem}). As a consequence, in order to ascertain the convergence of PM-C to global minimum of the problem~(\ref{eq:go_problem}) it is sufficient to prove the convergence the estimation $\omega^*$ to $f^*$ when $\varepsilon_1$ (search accuracy in~(\ref{eq:stop_criterion}) for the subproblem~(\ref{eq:subproblem_0})) tends to zero, i.e.,
\begin{equation}
  \label{eq:limit_omega}
  \lim_{\varepsilon_1 \rightarrow 0} \omega^* = f^*.
\end{equation}

If PM is applied for univariate minimization of subproblems~(\ref{eq:subproblem_i}), its parameters can differ at different levels of
recursion, in particular, one can take distinct accuracies in the termination criterion~(\ref{eq:stop_criterion}) and consider the vector
\begin{equation}
  \varepsilon = ( \varepsilon_1, \dots, \varepsilon_N ),
\end{equation}
%
where $\varepsilon_i$ is the termination accuracy during solving the subproblems~(\ref{eq:subproblem_i}) for the coordinate $y_i$, $1 \leq i \leq N$.

\begin{theorem}
  If the objective function $f(y)$ in the problem~(\ref{eq:go_problem}) satisfies the Lipschitz condition~(\ref{eq:lipschitz}) with constant $L > 0$, all the subproblems~(\ref{eq:subproblem_i}) are solved by PM with the parameter $m > L$ and accuracies $\varepsilon_i$ satisfy the inequalities
  \begin{equation}
    \label{eq:eps_next_lower_eps}
    \varepsilon_{i + 1} < \varepsilon_i, \; 1 \leq i < N,
  \end{equation}
  %
  then the assertion~(\ref{eq:limit_omega}) is true.
\end{theorem}

\begin{proof}
  Suppose that for a number $i$, $1 \leq i < N$, the inequality
  \begin{equation}
    \label{eq:f_omega_abs}
    | f^i(y_1, \dots, y_i) - \omega^i(y_1, \dots, y_i) | \leq \xi_i,
  \end{equation}
  %
  holds. This condition is met for $i = N$ with $\xi_N = 0$ as
  \begin{equation}
    \omega^N(y_1, \dots, y_N) = f^N(y_1, \dots, y_N) = f(y).
  \end{equation}

  As the corollary of the Lipschitz condition for the function $f(y)$, all the functions $f^i(y_1, \dots, y_i)$ are Lipschitzian as well with the same constant $L$ (see~\cite{bib51}). Then, in accordance with the Theorem~\ref{thm:theorem1} the inequality~(\ref{eq:f_omega_abs}) involves for the number $i - 1$ the estimation
  \begin{equation}
    | f^{i - 1}(y_1, \dots, y_{i - 1}) - \omega^{i - 1}(y_1, \dots, y_{i - 1}) | \leq \xi_i + \frac{L \varepsilon_i}{2},
  \end{equation}
  %
  and, as consequence, one can take
  \begin{equation}
    \xi_{i - 1} = \xi_i + \frac{L \varepsilon_i}{2}
  \end{equation}
  %
  Since the condition~(\ref{eq:f_omega_abs}) takes place for $i = N$ with $\xi_N = 0$ it is true for all $i$, $1 \leq i \leq N$, and we get the estimation
  \begin{equation}
    \label{eq:phi_omega_diff}
    | \varphi^* - \omega^* | \leq \frac{L}{2} \sum_{i = 1}^N \varepsilon_i,
  \end{equation}

  Then the statement~(\ref{eq:limit_omega}) results immediately from~(\ref{eq:eps_next_lower_eps}) and~(\ref{eq:phi_omega_diff}). The theorem has been proven.
\end{proof}

Finally, for the adaptive nested scheme with the Piyavskij's method inside (PM-A) its convergence follows from the convergence of PM-C by reasons which substantiated in~\cite{bib38} the convergence of GSA-A.

\section{Numerical experiments}
\label{sec:num_exp}
In this section we consider 7 global search algorithms and compare their efficiency experimentally. 6 methods combine the reduction schemes described above with two Lipschitzian one-dimensional methods and their generalizations to the H\"older case and were considered in previous section. To demonstrate the efficiency of these algorithms in comparison with a method of another nature we took the well-known method of Lipschitz optimization DIRECT~\cite{bib23,bib24} as well.

All the methods were compared on the test class of multiextremal functions GKLS from~\cite{bib48} being an easy-to-use source of functions with tuned complexity. The test sets consisting of 100 GKLS problems were taken for each of dimensions 2, 3, 4. For all the dimensions the hard class of GKLS was chosen with the following parameters:
\begin{itemize}
  \setlength\itemsep{0em}

  \item 10 local minima;
  \item radius of the attraction region of the global minimizer equal to $0.12$;
  \item standard function value $-1.0$ at the global minimizer;
  \item distance from the global minimizer to the vertex of the paraboloid equal to $0.9$.
\end{itemize}

The efficiency of algorithms was measured in accordance with the following two criteria: the average number $K$ of trials (evaluations of the objective function) executed during optimization of the functions from the test set and the number $P$ of test problems solved successfully. A test problem was supposed to have been solved if the trial with the lowest value obtained in the course of optimization was placed in a $\delta$-neighbourhood of the global minimizer.

In more details, the values $K$ and $P$ were estimated as follows. The algorithm with a given set of parameters solves $n$ optimization problems from the test class. After solving the $i$-th test problem we have the number $K_i$ of executed trials and the distance
\begin{equation}
  P_i = \sqrt{\sum_{j = 1}^{N} ( y_j^* - y_j^+ )^2}
\end{equation}
%
between the real global minimizer $y^* = (y_1^*, \dots, y_N^*)$ and the coordinate vector $y^+ = (y_1^+, \dots, y_N^+)$ corresponding to the lowest function value found by the algorithm.

After solving all the test problems the criteria $K$ and $P$ are calculated as
\begin{equation}
  K = \frac{1}{n} \sum_{i = 1}^{n} K_i, \quad P = \frac{1}{n}\sum_{i = 1}^{n} p_i,
\end{equation}
%
where for given accuracy $\delta > 0$
\begin{equation}
  p_i =
  \begin{cases}
    1, & P_i <    \delta, \\
    0, & P_i \geq \delta.
  \end{cases}
\end{equation}

Taking these criteria into consideration it is possible to apply the \emph{method of operational characteristics}~\cite{bib39,bib52} which allows one to present comparison results in a visual form. In accordance with this method an optimization algorithm solves problems of a test set several times using different values of its parameters and obtains a pair $(K, P)$ of the efficiency criteria for each variant of parameters chosen. The set of these pairs is called \emph{operational characteristic} of the algorithm. Plotting the operational characteristics on the plane $(K, P)$ enables to compare visually the efficiency of different optimization algorithms. Namely, if for the same $P$ the operational characteristic of an algorithm is located to the left of the characteristic of another method, the first algorithm is better because it has spent less number of trials $K$ for successful solving the same number of problems.

In experiments the algorithms on the base of the reduction schemes used in~(\ref{eq:m}) the identical parameter $r = 8$ providing for all 6 methods the sufficient  conditions of global convergence  and the curves of the operational characteristics for these algorithms and the method DIRECT were obtained by means of variation of parameters in stopping criteria of these methods.

The operational characteristics of the optimization algorithms underwent the examination on the 3-dimensional GKLS set are divided for better visualization into 3 groups and presented in Fig.~\ref{fig:ops_pm},~\ref{fig:ops_gsa},~\ref{fig:ops_gsa_pm}.

Fig.~\ref{fig:ops_pm} and Fig.~\ref{fig:ops_gsa} contain the operational characteristics for the algorithms of PM-type family and for the algorithms of GSA-type family correspondingly. Fig.~\ref{fig:ops_gsa_pm} compares the operational characteristics of the algorithms based on the adaptive nested scheme. In these figures the number of trials is plotted on the abscissa axis in the logarithmic scale.

As it follows from the results presented in Fig.~\ref{fig:ops_pm} and Fig.~\ref{fig:ops_gsa}, among the algorithms combining the methods of the same type with different schemes of dimensionality reduction the adaptive nested scheme is most efficient if it is necessary to solve problems with high reliability $P$. At the same time, the efficiencies of the reduction by means of Peano-type space-filling curves and the classical nested scheme differ slightly. Moreover, all 6 algorithms realizing the dimensionality reduction outperform significantly the method DIRECT.

With regard to comparison of the algorithms having demonstrated the best results in their families, the method GSA-A is some more efficient than PM-A.

The similar conclusions on the efficiency of the compared algorithms can be drawn after the experiments on the 2- and 4-dimensional GKLS classes. As a confirmation Tables~\ref{tab:trials_2d} and~\ref{tab:trials_4d} describe the results of testing for these dimensions. The left column contains the number $K$ of trials and the other columns correspond to the studied methods and show the quantity of problems solved successfully depending on the spent trials.

\section{Conclusion}
\label{sec:conclusion}
Numerical multidimensional global optimization algorithms based on several ideas of a dimensionality reduction have been considered. These algorithms reduce the multidimensional problem to one or several univariate subproblems solved by efficient methods of one-dimensional global optimization. Two approaches to reducing the dimension in multiextremal Lipschitzian optimization problems have been used: the reduction by means of Peano-type space-filling curves and the nested optimization scheme. In the framework of the second approach two versions of the nested reduction have been taken, namely, classical nested scheme and its recent generalization --- adaptive nested optimization. For solving the univariate subproblems the known algorithms proposed by Strongin and Piyavskij and the generalizations of these methods to the class of H\"olderian functions have been used.

For the considered algorithms the theoretical substantiation of convergence to global optimum has been given including both the known results and the new ones proved in the paper. For evaluation of efficiency a representative experimental testing of the methods has been carried out for different dimensions on sets of multiextremal functions belonging to the well-known GKLS test class being at present a classical tool for testing global optimization algorithms. Additionally, the popular method DIRECT has been tested to compare the reduction methods with an algorithm of the other nature. As a result of experiment, among all competitors the adaptive nested scheme has demonstrated the best efficiency.

The promising way of future researches consists in development of parallel algorithms based on the adaptive nested scheme, theoretical investigation of their properties and evaluation of parallelization efficiency, including comparison with existing parallel optimization methods.

\section*{Acknowledgements}
This research was supported by the Russian Science Foundation, project No. 15-11-30022 \enquote{Global optimization, supercomputing computations, and applications}.

\section*{References}
\bibliography{bibliography}

\appendix
\section{}

\begin{figure}[H]
  \centering
  \begin{tikzpicture}
    \node[root, fill = gray!20] {$\displaystyle\min_{y_1 \in [a_1, b_1]} f^1(y_1)$} [grow=right]
      child { node[level 1] (level_1) {$\displaystyle\min_{y_2 \in [a_2, b_2]} f^2(y_1^{k_1}, y_2)$}
        child { node[level 2] (level_2_1) {$\displaystyle\min_{y_3 \in [a_3, b_3]} f^2(y_1^{k_1}, y_2^{k_1, k_{2 k_1}}, y_3)$} }
        child { node[level 2] {$\displaystyle\min_{y_3 \in [a_3, b_3]} f^2(y_1^{k_1}, y_2^{k_1, 1}, y_3)$} }
      }
      child { node[level 1, fill = gray!20] {$\displaystyle\min_{y_2 \in [a_2, b_2]} f^2(y_1^2, y_2)$}
        child { node[level 2, fill = gray!20] (level_2_2) {$\displaystyle\min_{y_3 \in [a_3, b_3]} f^2(y_1^2, y_2^{2, k_{22}}, y_3)$} }
        child { node[level 2] {$\displaystyle\min_{y_3 \in [a_3, b_3]} f^2(y_1^2, y_2^{2, 1}, y_3)$} }
      }
      child { node[level 1] {$\displaystyle\min_{y_2 \in [a_2, b_2]} f^2(y_1^1, y_2)$}
        child { node[level 2] (level_2_3) {$\displaystyle\min_{y_3 \in [a_3, b_3]} f^2(y_1^1, y_2^{1, k_{21}}, y_3)$} }
        child { node[level 2] {$\displaystyle\min_{y_3 \in [a_3, b_3]} f^2(y_1^1, y_2^{1, 1}, y_3)$} }
      };
    \node[dots_large, above = 25pt of level_1] {$\dots$};
    \node[dots, above = 5pt of level_2_1] {$\dots$};
    \node[dots, above = 5pt of level_2_2] {$\dots$};
    \node[dots, above = 5pt of level_2_3] {$\dots$};
  \end{tikzpicture}
  \caption{\label{fig:nested_scheme} The subordination structure of the univariate subproblems~(\ref{eq:subproblem_i})}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{figure2.pdf}
  \caption{\label{fig:ops_pm} Operational characteristics of PM-type algorithms and DIRECT}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{figure3.pdf}
  \caption{\label{fig:ops_gsa} Operational characteristics of GSA-type algorithms and DIRECT}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{figure4.pdf}
  \caption{\label{fig:ops_gsa_pm} Operational characteristics of GSA-A and PM-A}
\end{figure}

\begin{center}
  \begin{longtable}{cccccccc}
    \caption{\label{tab:trials_4d} Number of problems solved successfully depending on the executed trials on four-dimensional class} \\
    \hline
    \multicolumn{1}{c}{K}      &
    \multicolumn{1}{c}{GSA-A}  &
    \multicolumn{1}{c}{GSA-C}  &
    \multicolumn{1}{c}{GGSA}   &
    \multicolumn{1}{c}{PM-A}   &
    \multicolumn{1}{c}{PM-C}   &
    \multicolumn{1}{c}{GPM}    &
    \multicolumn{1}{c}{DIRECT} \\
    \hline
    \endfirsthead

    \caption{Number of problems solved successfully depending on the executed trials on four-dimensional class (Continue)} \\
    \hline
    \multicolumn{1}{c}{K}      &
    \multicolumn{1}{c}{GSA-A}  &
    \multicolumn{1}{c}{GSA-C}  &
    \multicolumn{1}{c}{GGSA}   &
    \multicolumn{1}{c}{PM-A}   &
    \multicolumn{1}{c}{PM-C}   &
    \multicolumn{1}{c}{GPM}    &
    \multicolumn{1}{c}{DIRECT} \\
    \hline
    \endhead

    \hline
    \multicolumn{8}{r}{{Continued on the next page}} \\
    \hline
    \endfoot

    \hline
    \endlastfoot

    500    &   6 &   1 &   1 &   1 &   1 &   1 &   1 \\
    1000   &  22 &   1 &  20 &  21 &   1 &  15 &   1 \\
    2000   &  45 &   4 &  34 &  46 &   5 &  29 &   1 \\
    5000   &  82 &  48 &  53 &  79 &  47 &  57 &   1 \\
    10000  &  96 &  78 &  71 &  92 &  72 &  74 &   1 \\
    16000  & 100 &  87 &  79 & 100 &  86 &  83 &   1 \\
    30000  & 100 &  93 &  89 & 100 &  95 &  90 &  20 \\
    50000  & 100 & 100 & 100 & 100 & 100 & 100 &  50 \\
    70000  & 100 & 100 & 100 & 100 & 100 & 100 &  71 \\
    90000  & 100 & 100 & 100 & 100 & 100 & 100 &  81 \\
    120000 & 100 & 100 & 100 & 100 & 100 & 100 & 100 \\
  \end{longtable}
\end{center}

\begin{center}
  \begin{longtable}{cccccccc}
    \caption{\label{tab:trials_2d} Number of problems solved successfully depending on the executed trials on two-dimensional class} \\
    \hline
    \multicolumn{1}{c}{K}      &
    \multicolumn{1}{c}{GSA-A}  &
    \multicolumn{1}{c}{GSA-C}  &
    \multicolumn{1}{c}{GGSA}   &
    \multicolumn{1}{c}{PM-A}   &
    \multicolumn{1}{c}{PM-C}   &
    \multicolumn{1}{c}{GPM}    &
    \multicolumn{1}{c}{DIRECT} \\
    \hline
    \endfirsthead

    \caption{Number of problems solved successfully depending on the executed trials on two-dimensional class (Continue)} \\
    \hline
    \multicolumn{1}{c}{K}      &
    \multicolumn{1}{c}{GSA-A}  &
    \multicolumn{1}{c}{GSA-C}  &
    \multicolumn{1}{c}{GGSA}   &
    \multicolumn{1}{c}{PM-A}   &
    \multicolumn{1}{c}{PM-C}   &
    \multicolumn{1}{c}{GPM}    &
    \multicolumn{1}{c}{DIRECT} \\
    \hline
    \endhead

    \hline
    \multicolumn{8}{r}{{Continued on the next page}} \\
    \hline
    \endfoot

    \hline
    \endlastfoot

    10 &   1 &   1 &   8 &   1 &   1 &  18 &   3 \\
    20 &  35 &  32 &  53 &  30 &  26 &  52 &   6 \\
    30 &  82 &  71 &  80 &  68 &  59 &  71 &   9 \\
    40 &  96 &  94 &  91 &  92 &  86 &  84 &  13 \\
    50 & 100 &  99 &  97 & 100 &  96 &  92 &  16 \\
    80 & 100 & 100 & 100 & 100 & 100 & 100 &  24 \\
   150 & 100 & 100 & 100 & 100 & 100 & 100 &  35 \\
   250 & 100 & 100 & 100 & 100 & 100 & 100 &  47 \\
   500 & 100 & 100 & 100 & 100 & 100 & 100 &  80 \\
   800 & 100 & 100 & 100 & 100 & 100 & 100 & 100 \\
  \end{longtable}
\end{center}

\end{document}

### 2019-11-18 2020/PAVT_Nested_Full/pavt_paper.tex
Update Russian text

@@
 Общая схема организации вычислений с использованием нескольких процессов приведена на Fig.~\ref{MPI_tree}.
-% тут нужен наш традиционный рисунок - дерево процессов
+
+\begin{figure}
+\center{\includegraphics[width=1.0\linewidth]{MPI_Tree.pdf}}
+\caption{Дерево подзадача в параллельной схеме вложенной оптимизации}
+\label{MPI_tree}
+\end{figure}
+
 Процессы параллельной программы будут образовывать дерево. 
@@
 Каждая подзадача решается в отдельном процессе, обмен данными между процессами может быть организован с помощью MPI.
-Пересылки данных между процессами будут минимальные: требуется лишь передать от корня к потомкам координаты точек испытаний, и получить обратно значения функции в этих точках.
+Пересылки данных между процессами будут минимальные: требуется передать от корня к потомкам координаты точек $y^1, ..., y^p$, и получить обратно найденные минимальные значения
+\[
+\varphi(y^1) = \min_x f(x,y^1),\; ..., \; \varphi(y^p) = \min_x f(x,y^p).
+\]
+
 Подробное обсуждение алгоритмов, используемых на разных уровнях дерева процессов, приведено в следующих разделах.
@@
 Various modifications of this algorithm and the corresponding theory of
-convergence are given in \cite{Strongin2000,Barkalov2002,Gergel2005,Strongin2013}.
+convergence are given in \cite{Strongin2000,Sergeyev2013}.
 
@@
 
+Исследование масштабируемости предложенного параллельного алгоритма будем проводить, решая серию из 100 тестовых задач, в которых можно выделить параметры, влияющие локально.
+С целью имитации прикладных задач, обладающих указанным свойством, будем использовать функции вида
+\[
+f(x,y) = G(y)+R(x),
+\]
+где $G(y)$ -- многоэкстремальная функция размерности $N=2$, а $R(x)$ -- унимодальная функция размертности $N \gg 2$.
+В качестве многоэкстрмальной части задачи будем рассматривать функции 
+\begin{eqnarray} \nonumber \label{vagris}
+G(y)= -&\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2+\right. \\
+&\left.\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},\\ \nonumber
+\end{eqnarray}
+where
+\begin{eqnarray} \nonumber
+& y=(y_1,y_2)\in R^2, 0 \leq y_1,y_2 \leq 1, \\ \nonumber
+& g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2),  \\ \nonumber
+& h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2), \nonumber 
+\end{eqnarray}
+and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
+Данный класс функций часто используется для тестирования алгоритмов глобальной оптимизации.  
+В качестве локальной части задачи будем использовать модификацию функции Розенброка 
+\[
+R(x)= \sum_{i=1}^{N-1}{\left[(1-x_i)^2+100(x_{i+1}-x_i^2)^2\right]}, -2 \leq x_i \leq 2 , 1\leq i\leq N,
+\]
+размерности $N-2$, в которой точка экстремума смещается случайным образом в области поиска с помощью линейного преобразования координат.
+
+В качестве примера на Fig.~\ref{fig_level}(a),(b) приведены линии уровня одной пары функций $G(y)$ и $R(x)$.
+Данные функции являются сложными для соответствующих методов глобальной и локальной оптимизации, т.к.
+функции $G(y)$ являются существенно многоэкстремальными, а функции $R(x)$ обладают ярко выраженной овражной структурой.
+
+\begin{figure}
+\center
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{grishagin.png} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{grishagin.png} \\ (b)}
+\end{minipage}
+\caption{Линии уровня функций (a) $G(y)$ и (b) $R(x)$}
+\label{fig_level}
+\end{figure}
+
 \section{Conclusion}
+
+Результаты проведенных экспериментов на серии задач оптимизации, в которых можно выделить многоэкстремальную и унимодальную части задачи, показывают, что предложенная в работе схема организации параллельных вычислений обладает хорошим потенциалом параллелизма.
+В частности, при решении серии из 100 задач с числом переменных 52 (из которых две переменных оказывали глобальное влияние на целевую функцию, а пятьдесят влияли локально) получено линейное ускорение (в среднем).
+А решении задач с числом локально влияющих переменных $N=100$ было получено ускорение ...
+Проведенное исследование позволяет с уверенностью говорить о решении существенно многомерных задач оптимизации указанного класса.
+
+%Направлением дальнейших исследований может являться как расширение числа используемых локальных и глобальных методов, так и расширения круга рассматриваемых задач, включая задачи оптимизации с ограничениями, а также задачи многокритериальной оптимизации.
 

### 2019-11-19 2019/PACT_MCO/source/2019_pavt.tex
Добавил исходники статей по ГО за 2019г. (Гергель Козинов)

--- 2018-12-11 2019/PACT_MCO/2019_pavt.tex
@@
 In the present paper, the results of the performed investigations on the generalization of the decision making problem statements \cite{x11,x12} and on the development of the highly efficient global optimization methods utilizing the whole search information obtained in the course of computations \cite{x13,x14,x15,x16,x17} are presented.
+Parallel algorithms developed earlier are presented in papers \cite{x14,x15,x16,x17}.
+The proposed algorithms are able to efficiently use hundreds and thousands of computational cores \cite{x17_1}.
+In the scope of this paper, a comparative analysis of the implemented parallel algorithms on modern computing systems is carried out.
+The implementations of algorithms for shared memory, distributed memory and heterogeneous computing systems are compared.
+As a result of a comparative analysis, an optimal running configuration is selected.
 
@@
 The values of characteristics $w(y)$ are supposed to be non-negative, and the decreasing of these ones corresponds to the increasing of the efficiency of the chosen variants.
-Also, it is supposed that the characteristics $w_j (y)$, $1 \leq j \leq M$ can be multiextremal and the computing of the values of these ones can require a large amount of computations.
-Besides, within the framework of considered approach, the characteristics $w_j(y)$, $1 \leq j \leq M$ are supposed to satisfy the Lipschitz condition
+Also, it is supposed that the characteristics $w_j (y)$, $1 \leq j \leq M$, can be multiextremal and the computing of the values of these ones can require a large amount of computations.
+Besides, within the framework of considered approach, the characteristics $w_j(y)$, $1 \leq j \leq M$, are supposed to satisfy the Lipschitz condition
 \begin{equation}
@@
 where $L_j$ is the Lipschitz constant for the characteristic $w_j(y)$, $1 \leq j \leq M$, and  $\|*\|$ denotes the Euclidean norm in $R^N$.
-It is important to note that the fulfillment of the Lipschitz condition corresponds to the practical applications -- at small variations of the parameters $y \in D$ the corresponding variations of the values of characteristics $w_j(y)$, $1 \leq j \leq M$ are limited as a rule.
+It is important to note that the fulfillment of the Lipschitz condition corresponds to the practical applications -- at small variations of the parameters $y \in D$ the corresponding variations of the values of characteristics $w_j(y)$, $1 \leq j \leq M$, are limited as a rule.
 
@@
 The requirements to the optimality of the chosen variants $y \in D$ of the DMP can be defined in the following way.
-First of all, the characteristics $w_j(y)$, $1 \leq j \leq M$ should be selected, for which the achievement of the minimum possible values is necessary.
+First of all, the characteristics $w_j(y)$, $1 \leq j \leq M$, should be selected, for which the achievement of the minimum possible values is necessary.
 Defining a set of indices $F=\{ i_1,i_2, \dots ,i_s \}$ of such characteristics defines the vector criterion of efficiency
@@
 
-For the characteristics $w_j(y)$, $1 \leq j \leq M$ not included into the vector criterion of efficiency with the set of indices $G=\{ j_1,j_2, \dots ,j_m\}$ the sufficient levels of efficiency should be defined by a tolerance vector $q=(q_1,q_2,\dots,q_m)$.
+For the characteristics $w_j(y)$, $1 \leq j \leq M$, not included into the vector criterion of efficiency with the set of indices $G=\{ j_1,j_2, \dots ,j_m\}$ the sufficient levels of efficiency should be defined by a tolerance vector $q=(q_1,q_2,\dots,q_m)$.
 The availability of tolerances allows defining a vector function of constraints
@@
 Along with the usefulness of such an approach (all effective decisions of the MCO problem are found), the use of these methods appears to be difficult often because of high computational complexity of obtaining the estimates of the Pareto set.
-Besides, the estimate of the whole set $PD(Q)$ may appear tо be redundant when to solve the MCO problem it is enough to obtain several particular effective decisions (that takes place in the practical applications often).
+Besides, the estimate of the whole set $PD(Q)$ may appear to be redundant when to solve the MCO problem it is enough to obtain several particular effective decisions (that takes place in the practical applications often).
 As a result, more widely used approach to solving the MCO problems is based on the scalarization of the vector criterion into some general scalar criterion of efficiency, the optimization of which can be performed with the use of already existing optimization methods.
@@
 
-The general approach used in the present work consists in the reduction of solving a MMGO problem to the solving of a series of single-criterion global optimization problems \footnote{Due to initial assumptions on possible multiextremality of the characteristics $w_j(y), 1 \leq j \leq M$ from (\ref{eq:01})}:
+The general approach used in the present work consists in the reduction of solving a MMGO problem to the solving of a series of single-criterion global optimization problems \footnote{Due to initial assumptions on possible multiextremality of the characteristics $w_j(y), 1 \leq j \leq M$, from (\ref{eq:01})}:
 \begin{equation}
@@
 \end{equation}
-where $g(y)$ is the vector function of constraints from (\ref{eq:06}), and $F(\lambda,y)$ is the minimized convolution of the partial criteria $f_i(y(x))$, $1 \leq i \leq s$ of the MCO problem (\ref{eq:05})
+where $g(y)$ is the vector function of constraints from (\ref{eq:06}), and $F(\lambda,y)$ is the minimized convolution of the partial criteria $f_i(y(x))$, $1 \leq i \leq s$, of the MCO problem (\ref{eq:05})
 \begin{equation}
@@
 
-The numerical solving of the global optimization problems is reduced usually to a sequential computing the values of characteristics $w(y)$ at the points $y^i$, $0\leq i \leq k$ of the search domain $D$ \cite{x18,x22,x23,x24,x25}.
+The numerical solving of the global optimization problems is reduced usually to a sequential computing the values of characteristics $w(y)$ at the points $y^i$, $0\leq i \leq k$, of the search domain $D$ \cite{x18,x22,x23,x24,x25}.
 The data obtained as a result of computations can be represented in a form of matrix of search information (MSI)
@@
 \begin{itemize}
-	\item $x_i$, $1 \leq i \leq k$ are the reduced points of the executed global search iterations where the criteria values have been computed; 
-	\item $z_i$, $g_i$, $1 \leq i \leq k$ are the values of the scalar criterion and of the constraints at the points $x_i$, $1 \leq i \leq k$ for current optimization problem $F(\lambda,y(x))$ from (\ref{eq:15}) being solved i.
+	\item $x_i$, $1 \leq i \leq k$, are the reduced points of the executed global search iterations where the criteria values have been computed; 
+	\item $z_i$, $g_i$, $1 \leq i \leq k$, are the values of the scalar criterion and of the constraints at the points $x_i$, $1 \leq i \leq k$, for current optimization problem $F(\lambda,y(x))$ from (\ref{eq:15}) being solved i.
 e.
@@
 \end{equation}
-	\item $l_i$, $1 \leq i \leq k$ are the indices of the global search iterations, at which the points $x_i$, $1 \leq i \leq k$ have been generated, these indices are used to storing the correspondence between the reduced points and the multidimensional ones at the executed iterations i.
+	\item $l_i$, $1 \leq i \leq k$, are the indices of the global search iterations, at which the points $x_i$, $1 \leq i \leq k$, have been generated, these indices are used to storing the correspondence between the reduced points and the multidimensional ones at the executed iterations i.
 e.
@@
 The matrix of the search state $A_k$ contains the search information transformed to current scalar reduced problem (\ref{eq:15}) being solved.
-Besides, the search information in the MSS is arranged according to the values of coordinates of the points $x_i$, $1 \leq i \leq k$ for more efficient execution of the global search algorithms; the arranged representation of the points is reflected by the use of the lower index i.
+Besides, the search information in the MSS is arranged according to the values of coordinates of the points $x_i$, $1 \leq i \leq k$, for more efficient execution of the global search algorithms; the arranged representation of the points is reflected by the use of the lower index i.
 e.
@@
 \label{eq:20}
-y^{k+1}=Y(\Omega_k),k=1,2,\dots.
+y^{k+1}=Y(\Omega_k),k=1,2,\dots
 \end{equation}
-where $Y$ is the rule for computing the points $y^i$, $0 \leq i \leq k$ of the applied optimization algorithm.
+where $Y$ is the rule for computing the points $y^i$, $0 \leq i \leq k$, of the applied optimization algorithm.
 Such an adaptive choice of the points of executed search iterations can accelerate the finding of the effective decisions.
@@
 
-It should be noted that the availability of MSS $\Omega_k$ from (\ref{eq:18}) allows reducing the results of all preceding computations $z_i$,  $1 \leq i \leq k$ stored in the MSS $A_k$ to the values of current optimization problem $F(\lambda,y(x))$ from (\ref{eq:15}) being solved for any new parameter of the MCO problem statement $P$ from (\ref{eq:09}) and for any new value of the convolution coefficients $\lambda$ from (\ref{eq:14}) on the basis of the values of characteristics $w^i$, $1 \leq i \leq k$ computed earlier and stored in $\Omega_k$ without any repeated time-consuming computation of the values of $w(y)$ from (\ref{eq:01}) i.e.
+It should be noted that the availability of MSS $\Omega_k$ from (\ref{eq:18}) allows reducing the results of all preceding computations $z_i$,  $1 \leq i \leq k$, stored in the MSS $A_k$ to the values of current optimization problem $F(\lambda,y(x))$ from (\ref{eq:15}) being solved for any new parameter of the MCO problem statement $P$ from (\ref{eq:09}) and for any new value of the convolution coefficients $\lambda$ from (\ref{eq:14}) on the basis of the values of characteristics $w^i$, $1 \leq i \leq k$, computed earlier and stored in $\Omega_k$ without any repeated time-consuming computation of the values of $w(y)$ from (\ref{eq:01}) i.e.
 \begin{equation}
@@
 
-The method of organizing the parallel computations considered above can be developed further by utilizing the information compatibility of the problems from the set $\mathbb{F}$ -- as has been noted in Section {\ref{sec:04}}, the values of characteristics $w^i$, $1 \leq i \leq k$ computed when solving any problem $F_1(\lambda,y(x))$ can be transformed to the values of any other problem $F_2(\lambda,y(x))$.
+The method of organizing the parallel computations considered above can be developed further by utilizing the information compatibility of the problems from the set $\mathbb{F}$ -- as has been noted in Section {\ref{sec:04}}, the values of characteristics $w^i$, $1 \leq i \leq k$, computed when solving any problem $F_1(\lambda,y(x))$ can be transformed to the values of any other problem $F_2(\lambda,y(x))$.
 This result allows proposing the following general scheme of parallel computations for simultaneous solving the problems from the set $\mathbb{F}$.
@@
 \begin{enumerate}
-	\item The choice of $q$, $q \geq 1$ points $x^i$, $1\leq i\leq q$ of the next global search iteration is performed according to the rule (\ref{eq:20}) of the optimization method; the choice of the points is performed taking into account the available search information $\Omega_k$ and $A_k$; the number of generated points $q$, $q\geq 1$ is determined by the number of computational cores employed for solving the problem $F(\lambda,y(x))$;
-	\item For each chosen one-dimensional point $x^i \in [0,1]$, $1 \leq i \leq q$ of current global search iteration, the multidimensional image $y^i \in D$, $1 \leq i \leq q$ is determined according to the mapping $y(x)$.
-Then, each computed image $y^i \in D$, $1 \leq i \leq q$ is transmitted to all employed computational nodes to avoid the multiple choice of the same points in the domain $D$ when solving the problems from the set $\mathbb{F}$;
+	\item The choice of $q$, $q \geq 1$ points $x^i$, $1\leq i\leq q$, of the next global search iteration is performed according to the rule (\ref{eq:20}) of the optimization method; the choice of the points is performed taking into account the available search information $\Omega_k$ and $A_k$; the number of generated points $q$, $q\geq 1$ is determined by the number of computational cores employed for solving the problem $F(\lambda,y(x))$;
+	\item For each chosen one-dimensional point $x^i \in [0,1]$, $1 \leq i \leq q$, of current global search iteration, the multidimensional image $y^i \in D$, $1 \leq i \leq q$, is determined according to the mapping $y(x)$.
+Then, each computed image $y^i \in D$, $1 \leq i \leq q$, is transmitted to all employed computational nodes to avoid the multiple choice of the same points in the domain $D$ when solving the problems from the set $\mathbb{F}$;
 	\item The values of characteristics $w(y)$ from (\ref{eq:01}) are computed at all points $y^i \in D$, $1 \leq i \leq q$; for each particular point $y^i \in D$, $1 \leq i \leq q$, these computations are performed in parallel using different computational cores; the computed values of characteristics $w(y)$ are transmitted to all employed computational nodes to include the obtained data into the search information $\Omega_k$ and $A_k$.
@@
 
-According to the presented computational scheme of parallel computations, each computational unit would contain the identical copies of the matrix of search information $\Omega_k$ from (\ref{eq:18}); the matrices $A_k$ from (\ref{eq:19}) would contain different sets of the global search points $x_i$, $1\leq i\leq k$ because of the use of different evolvents $y(x)$ for the dimensionality reduction and of different values of the scalar criterion and constraints $z_i$, $g_i$, $1\leq i\leq k$ corresponding to the problems $F(\lambda,y(x))$ from the set $\mathbb{F}$ being solved.
+According to the presented computational scheme of parallel computations, each computational unit would contain the identical copies of the matrix of search information $\Omega_k$ from (\ref{eq:18}); the matrices $A_k$ from (\ref{eq:19}) would contain different sets of the global search points $x_i$, $1\leq i\leq k$, because of the use of different evolvents $y(x)$ for the dimensionality reduction and of different values of the scalar criterion and constraints $z_i$, $g_i$, $1\leq i\leq k$, corresponding to the problems $F(\lambda,y(x))$ from the set $\mathbb{F}$ being solved.
 
@@
 The developed general computational scheme includes the concurrent parallel computational methods for the computational systems with the shared memory, the distributed parallel computations using the multiple mappings for the dimensionality reduction, and the multilevel nesting of the parallel computations for the high-performance computational systems.
-In general, the proposed computational scheme ensures the opportunity of an efficient application of the high-performance computational systems with a large number (tens and hundreds thousand) cores/processors for solving the complex global optimization problems.
+In general, the proposed computational scheme can provide the opportunity of an efficient application of the high-performance computational systems with a large number cores/processors for solving the complex global optimization problems.
 Besides, this general scheme can be used for the organization of the parallel computations for a wide range of algorithms for solving the time-consuming problems of decision making (in particular, for the information-statistical global optimization algorithms).
@@
 \doi{10.1007/978-3-319-71255-0\_10}
+\bibitem {x17_1}Gergel, V., Kozinov, E.: GPU-based computations in multicritical optimization .
+Communications in Computer Science and Information Science, \textbf{965}, 88--100 (2019).
+\doi{10.1007/978-3-030-05807-4\_8}
 \bibitem {x18}Strongin, R., Sergeyev, Ya.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2nd ed. 2013, 3rd ed. 2014).

### 2019-11-19 2019/PPAM_Gergel_Kozinov/PPAM_2019_10_29.tex
Добавил исходники статей по ГО за 2019г. (Гергель Козинов)

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{mathtools,amssymb}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Parallel Computations for Various Scalarization Schemes in Multicriteria Optimization Problems\thanks{This research was supported by the Russian Science Foundation, project No 16-11-10150 ''Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance.''}}
%
\titlerunning{Parallel Computations in Multicriteria Optimization}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Victor Gergel\orcidID{0000-0002-4013-2329} \and Evgeny Kozinov \orcidID{0000-0001-6776-0096}}

\authorrunning{V. Gergel et al.} % if too long for running head
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
              \email{gergel@unn.ru, evgeny.kozinov@itmm.unn.ru}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In the present paper, a novel approach to parallel computations for solving time-consuming multicriteria global optimization problems is presented. This approach includes various methods for the scalarization of vector criteria, dimensionality reduction with the use of the Peano space-filling curves, and efficient global search algorithms. The applied criteria scalarization methods can be altered in the course of computations in order to better meet the stated optimality requirements. To reduce the computational complexity of multicriteria problems, the methods  developed feature an extensive use of all the computed optimization information and are well parallelized for effective performance on high-performance computing systems. Numerical experiments confirmed the efficiency of the developed approach.

\keywords{Multicriteria optimization \and criteria scalarization \and global optimization \and parallel computations.}
\end{abstract}
%
%
%
\section{Introduction} \label{sec:1}

Multicriteria optimization (MCO) problems arise in various applications of science and technology. The wide scope of application of MCO problems determines the intensity of research in the field \cite{c1,c2,c3,c4,c5}.

Usually, the criteria of the MCO problems are contradictory, and obtaining the decisions, which provide the best values with respect to all criteria simultaneously, is impossible. In such situations, a set of efficient (non-dominated) decisions is considered for such MCO problems,  where the improvement of
the values of some criteria cannot be achieved without sacrificing the efficiency values with respect to other criteria. When solving MCO problems, it may become necessary to obtain the whole set of the efficient decisions (the Pareto set) that may require performing a large amount of computations. Besides, the analysis of a large set of the efficient decisions may appear to be
difficult for a decision maker. Therefore, in practice the finding of only a relatively small set of efficient decisions may be justified. Constructing such a limited set of efficient decisions is performed usually according to the optimality requirements defined by the decision-maker.

The restricted set of the computed efficient decisions leads to a notable reduction in the amount of required computations. However, the criteria of efficiency may have a complex multiextremal form, and the computation of the criteria values may appear to be time-consuming. One way to overcome the considerable computational complexity when solving MCO problems is to use high-performance supercomputing systems.

In the present paper, the research results on the development of highly efficient parallel methods of the multicriteria optimization using the search information obtained in the course of computations \cite{c6,c7,c8,c9,c16} are presented. A novel contribution consists in the development of an approach whereby it is possible to use different MCO problem criteria scalarization schemes taking into account different optimality requirements to desirable decisions. For the scalarization schemes considered in this study, a general method of parallel computations was proposed and the computational evaluation of the developed approach's efficiency was performed.

The further structure of the paper is as follows. In Section 2, the statement of the multicriteria optimization problems is given. In Section \ref{sec:3}, a general scalarization scheme for the MCO problem criteria involving various kinds of criteria convolution is proposed. In Section \ref{sec:4}, the search information obtained in the course of computations, which can be reused after the altering of the applied scalarization schemes, is considered. Also, this section presents a general scheme for the parallel execution of the global search algorithms, in the framework of which an efficient parallel method for solving time-consuming global optimization problems is considered. Section \ref{sec:5} contains the results of the numerical experiments confirming that the approach developed by the authors is quite promising. In the Conclusions section, the obtained results are discussed and possible main areas for further studies are outlined.

\section{Multicriteria optimization problem statement} \label{sec:2}

In the most general form, the multicriteria optimization problem is defined in the following way 
\begin{equation}
f(y) \to min, y \in D :g(y) \leq 0 ,
\label{eq:1}
\end{equation}
where $y=(y_1,y_2, \dots ,y_N)$ is a vector of varied parameters, \\$f(y)=(f_1 (y),f_2 (y), \dots ,f_s (y))$ is the vector of the efficiency criteria, $D \subset R^N$ is the search domain 
\begin{equation}
D = \{ y\in R^N:a_i \leq y_i \leq b_i, 1 \leq i \leq N \},
\label{eq:2}
\end{equation}
for given vectors $a$ and $b$, and $g(y)= (g_1 (y),g_2 (y), \dots ,g_m (y))$ are the constraints of the MCO problem (the conditions of feasibility of the chosen decisions $y \in D$).

In the most complex cases, the criteria $f_i (y)$, $1 \leq i \leq s$, can be multiextremal and computing the values of these ones can appear to be time-consuming. Usually, the criteria $f_i (y)$, $1 \leq i \leq s$, satisfy the Lipschitz condition 
\begin{equation}
|f_i (y_1 )-f_i (y_2 ) | \leq L_i \|y_1-y_2\|,1 \leq i \leq s,
\label{eq:3}
\end{equation}
where $L_i$ are the Lipschitz constants for the criteria $f_j (y)$, $1 \leq j \leq s$ and $\|*\|$ denotes the Euclidean norm in $R^N$.

Without any loss in generality, in further consideration we will assume the criteria $f_i (y)$, $1 \leq i \leq s$ to be non-negative and their decrease corresponds to the increase in the efficiency of the chosen decisions.

\section{Reducing multicriteria optimization to multiple
global search problems} \label{sec:3}

As mentioned above, for solving MCO problems it is required to find a set of the efficient (Pareto optimal) decisions, for which the values cannot be improved with respect to all criteria simultaneously. However, a large amount of computations may be required to  find a complete set of efficient decisions. A commonly used approach to reducing the computational costs consists in finding of only a relatively small set of efficient decisions, defined according to the optimality requirements of the decision maker. In many cases, the required decisions are found by means of reducing the vector criteria to some general scalar efficiency function that allows applying a variety of already existing global optimization methods for solving MCO problems. Possible scalarization methods include, for example, the weighted sum method, the compromise programming method, the weighted min-max method, as well as many other methods -- see, for example, \cite{c2,c4}. Within the developed approach, it was proposed to employ the following methods of scalarization of the vector criterion of MCO problems.

1. One of the widely applied scalarization methods consists in the use of the min-max convolution of the criteria \cite{c4,c6}:
\begin{equation}
\begin{split}
F_1 (\lambda,y)=\max{(\lambda_i f_i (y),1 \leq i \leq s)}, \\
\lambda=(\lambda_1,\lambda_2,\dots,\lambda_s)\in \Lambda \subset R^s: \sum_{i=1}^s{\lambda_i=1}, \lambda_i \geq 0, 1 \leq i \leq s.
\end{split}
\label{eq:4}
\end{equation}

2. In the case when the criteria can be arranged in the order of importance, the method of successive concessions (MSC) is usually applied \cite{c5,c6}. In this method, the multistage computations can be reduced to solving the following scalar optimization problem 
\begin{equation}
F_2(\delta,y)=f_s(y), f_i(y) \leq f_i^{min}+\delta_i (f_i^{max}- f_i^{min}), 1 \leq i < s, g(y) \leq 0,y \in D,
\label{eq:5}
\end{equation}
where $f_i^{min}$, $f_i^{max}$, $1 \leq i < s$, are the minimum and the maximum values of the criteria in the domain $D$ from (\ref{eq:2}), and $0 \leq \delta_i \leq 1$, $1 \leq i < s$, are the concessions with respect to each criterion.

3. If an estimate of the criteria values for the required decision exists a \textit{priori} (for example, based on some ideal decision), the solution of a MCO problem can consist in finding an efficient decision that matches the predefined optimality values \cite{c2,c4}:
\begin{equation}
F_3 (\theta,y)= 1/s \sum_{i=1}^s{\theta_i(f_i(y)-f_i^*)^2 }, g(y) \leq 0, y \in D,
\label{eq:6}
\end{equation}
where $F_3 (\theta,y)$ is the mean square deviation of the decision $y \in D$ from the sought ideal decision $y^*$, and the values $0\leq \theta_i \leq 1$, $1 \leq i < s$, are the importance parameters of the approximation accuracy with respect to each variable $y_i$, $1 \leq i \leq N$.

In the general case, the statements of the global optimization problems generated by the criteria scalarization schemes (\ref{eq:4})-(\ref{eq:6}) can be represented in the form:
\begin{equation}
  \min\varphi(y)=\min{F(\alpha,y)},  g(y) \leq 0, y \in D, 
\label{eq:7}
\end{equation}
where $F$ is the objective function generated as a result of the criteria scalarization, $\alpha$ is the vector of parameters of the criteria convolution, $g(y)$ are the constraints of the MCO problem, and $D$ is the search domain from (\ref{eq:2}). It should be noted that due to the possibility of altering the optimality requirements in the course of computations, the form of the function $\varphi(y)$ from (\ref{eq:7}) can vary. Thus, it may be necessary to alter the scalarization method employed (\ref{eq:4})-(\ref{eq:6}) and/or to change the convolution parameters $\lambda$, $\delta$, and $\theta$. Such variations form a set of scalar global optimization problems
\begin{equation}
  \mathbb{F}_T=\{ F_k (\alpha_i,y): 1\leq i \leq T, k=1,2,3 \}.
\label{eq:9}
\end{equation}

This set of problems can be formed sequentially in the course of computations. The problems in this set can be solved either sequentially or simultaneously -- in the time sharing mode or in parallel with the use of high performance computational systems. The opportunity of forming the set $\mathbb{F}_T$ allows formulating \textit{a new class of multicriteria global optimization problems}.

Within the framework of the  proposed approach, one more step of transformation of the problems $F(\alpha,y)$ from the set $\mathbb{F}_T$ is applied -- namely, the dimensionality reduction with the use of the Peano space-filling curves (evolvents) $y(x)$ providing an unambiguous mapping of the interval $[0,1]$ onto an $N$-dimensional hypercube $D$ \cite{c10,c11}. As a result of such reduction, a multidimensional problem (\ref{eq:7}) is reduced to a one-dimensional global optimization problem:

\begin{equation}
F(\alpha,y(x^*)) = \min{\{F(\alpha,y(x)): g(y(x))\leq 0, x \in [0,1]\}}.
\label{eq:10}
\end{equation}

The dimensionality reduction allows applying many highly efficient\\ one-dimensional global search algorithms -- see, for example, \cite{c10,c11,c12,c14} -- for solving the problems of the set $\mathbb{F}_T$ from (\ref{eq:9}). It should be also noted that according to (\ref{eq:3}) the function $F(\alpha,y(x))$ satisfies the uniform H\"{o}lder condition with a constant $H$ \cite{c10,c11} i.e.
\begin{equation}
  |F(\alpha,y(x' )-F(\alpha,y(x'')|\leq H |x'-x''|^{1/N}.
\label{eq:11}
\end{equation}

\section{Parallel computations for solving multiple global optimization problems} \label{sec:4}

The numerical solving of the MCO problems consists usually in sequential computing of the values of the criteria $f(y)$ and the constraints $g(y)$ from (\ref{eq:1}) at the points $y^i$, $0 \leq i \leq k$, of the search domain $D$ \cite{c10,c12,c13,c14}. After the scalarization of the vector criterion (\ref{eq:7}) and the application of the dimensionality reduction (\ref{eq:10}), the obtained search information takes the form of the set:
\begin{equation}
  A_k=\{(x_i,z_i,f_i,g_i )^T:1 \leq i \leq k\},
\label{eq:12}
\end{equation}
where $x_i$, $1 \leq i \leq k$ are the reduced points of the executed global search iterations arranged in ascending order of the coordinates and $z_i$,  $f_i$, $g_i$, $1 \leq i \leq k$, are the values of the scalar criterion, the objective function criteria and the constraints respectively at these points for the current optimization problem $F(\alpha,y(x))$.

The availability of the set $A_k$ from (\ref{eq:12}) allows the transformation of the values of the criteria and constraints calculated earlier into the values of the current optimization problem $F(\alpha,y(x))$ from (\ref{eq:7}) without repeating the time-consuming computations of the values of criteria and constraints i. e.
\begin{equation}
  (x_i,f_i,g_i ) \to z_i = F(\alpha,y(x_i)) ,1 \leq i \leq k.
\label{eq:13}
\end{equation}

In this way, the whole search information $A_k$ from (\ref{eq:12}) recalculated according to (\ref{eq:13}) can be reused to continue the solution of the MCO problem. This can provide a considerable reduction of the amount of computations performed for solving every next MCO problem of the set $\mathbb{F}_T$ from (\ref{eq:9}) up to a limited number of global search iterations.

A further increase in computational efficiency when solving MCO problems can be provided by means of parallel computations. Within the framework of the developed approach, a general parallel computation scheme for solving global optimization problems is applied -- namely, parallel computations are provided by the simultaneous computing of the values of the minimized function $F(\alpha,y(x))$ from (\ref{eq:7}) at several different points of the search domain $D$ \cite{c6,c10}. Such an approach provides the parallelization of the most time-consuming part of the global search computations and can be applied to any global optimization method for a variety of global optimization problems.

The state-of-the-art in the field of global optimization is reviewed, for example, in \cite{c10,c12,c13}. In the present paper, the proposed approach is based on the information-statistical theory of global search \cite{c10}. This theory has been applied as a basis for the development of a large number of efficient methods of multiextremal optimization -- see, for example, \cite{c6,c7,c8,c9,c10,c11,c15,c16,c17,c18}. The main ideas of the algorithms developed consist in the dimensionality reduction with the use of the Peano space-filling curves and in parallel computations by simultaneous computing the values of the minimized functions at several points of the search domain. From the general point of view, the developed \textit{multidimensional parallel global search algorithms} (MPGSA) can be presented by a unified computation scheme as follows \cite{c6,c10,c15}.

Let $p>1$ be the number of the applied processors (cores) with shared memory. At the initial iteration of the MGGSA, the computing of the minimized function values at $p$ arbitrary points of the interval $(0,1)$ is performed (the obtaining of the function value will be called hereafter \textit{a trial}). Then, let us assume $\tau$, $\tau > 1$ global search iterations to be completed. The choice of the trial points for the next $(\tau+1)^{th}$ iteration is determined by the following rules.

\textit{Rule 1.} Renumber the points of the trial points by the lower indices in the order of increasing values of coordinate\footnote{It should be noted that the condition (\ref{eq:14}) determines the need to order the search information stored in the set $A_k$ from (\ref{eq:12}).}
\begin{equation}
0 = x_0 \leq x_1 \leq \dots \leq x_i \leq \dots \leq x_k \leq x_{k+1}=1
\label{eq:14}
\end{equation}

The points $x_0$, $x_{k+1}$ were introduced additionally for the convenience of further presentation, and $k=\tau p$ is the total number of trials performed earlier.

\textit{Rule 2.} For each interval $(x_{i-1} ,x_i)$, $1 \leq i \leq k+1$, compute the quantity $R(i)$ called hereafter the \textit{characteristic} of the interval.

\textit{Rule 3.} Arrange the characteristics of the intervals in the decreasing order 
\begin{equation}
 R(t_1 ) \geq R(t_2 )\geq \dots \geq R(t_{k-1} ) \geq R(t_k)
\label{eq:15}
\end{equation}

Select $p$ intervals with the indices $t_j$, $1 \leq j \leq p$, having the highest characteristics.

\textit{Rule 4.} Perform new trials at the points of the interval $(0,1)$
\begin{equation}
 x^{k+1}, x^{k+2},\dots,x^{k+p},
\label{eq:16}
\end{equation}
placed in the intervals with the highest characteristics from (\ref{eq:15}).

The stopping condition, according to which the trials are terminated, is defined as 
\begin{equation}
 (x_t-x_{t-1})^{1/N} \leq \varepsilon,
\label{eq:17}
\end{equation}
fulfilled for at least one of the indices $t_j$, $1 \leq j \leq p$, from (\ref{eq:15}), $N$ is the dimensionality, and $\varepsilon > 0$ is the predefined accuracy of solving the optimization problem. If the stopping condition is not met, the iteration index $\tau$ is incremented by 1, and a new global search iteration is performed.

According to this general scheme, particular global search algorithms are distinguished by expressions applied for computing the characteristics $R(i)$, $1 \leq i \leq k+1$, from (\ref{eq:15}) and the points of the next trials $x^{k+j}$, $1 \leq j \leq p$, from (\ref{eq:16}) in the intervals with the highest characteristics. Thus, for example, for the multidimensional generalized global search algorithm  (MGGSA) \cite{c10}, the interval characteristic takes the form 
\begin{equation}
 R(i) = m \varrho_i + \frac{(z_i-z_{i-1})^2}{m \varrho_i} - 2 (z_i+z_{i-1}), 1 \leq i \leq k+1 ,
\label{eq:18}
\end{equation}
where $m$ is the numerical estimate of the H\"{o}lder constant from (\ref{eq:11}), $z_i$, $1 \leq i \leq k+1$, are the computed values of the minimized function $F(\alpha,y(x))$ from (\ref{eq:10}) at the points of the performed global search iterations, and $\varrho_i = (x_i - x_{i-1})^{1/N}$, $1 \leq i \leq k+1$.

The MGGSA rule for computing the next trial point $x^k \in (0,1)$ is defined as
\begin{equation}
 x^{k+1} = \frac{x_t + x_{t-1}}{2}-sign(z_t-z_{t-1})\frac{1}{2r}\bigl[\frac{|z_t-z_{t-1}|}{m}\bigr]^N, 
\label{eq:19}
\end{equation}
where $r$, $r>1$ is a parameter of the MGGSA algorithm.

To clarify the computational scheme considered, the following can be noted. The computed characteristics $R(i)$, $1 \leq i \leq k+1$, from (\ref{eq:15}) can be considered as some measures of importance of the intervals from the viewpoint of their containing the global minimum point. Thus, the scheme of selecting the interval for executing the new trial becomes clear -- the point of every next trial $x^{k+1}$ from (\ref{eq:19}) is selected in the interval with the highest value of the interval characteristic (i. e. in the interval where the global minimum is most likely to be found).

The conditions of convergence for the algorithms developed in the framework of the information-statistical theory of global search and the conditions of non-redundancy of parallel computations were considered in \cite{c10}.

\section{Results of numerical experiments} \label{sec:5}

Numerical experiments were carried out on the Lobachevsky supercomputer at the State University of Nizhny Novgorod (the operating system -- CentOS 6.4, the management system -- SLURM). Each supercomputer node had 2 Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM processors. Each processor had 8 cores (i. e. a total of 16 CPU cores were available on each node).

In these experiments, bi-criteria two-dimensional MCO problems (i. e. $N=2$, $s=2$) were solved. Multiextremal functions defined by the relations \cite{c10}:
\begin{equation}
\begin{split}
f(y_1, y_2) = - \left\{ AB + CD \right\}^{1/2} \\
AB = \Big(\sum_{i=1}^7 \sum_{j=1}^7 [A_{ij} a_{ij}(y_1, y_2) + B_{ij} b_{ij}(y_1, y_2)]\Big)^2 \\
CD = \Big(\sum_{i=1}^7 \sum_{j=1}^7 [C_{ij} a_{ij}(y_1, y_2) - D_{ij} b_{ij}(y_1, y_2)]\Big)^2 
\end{split}
\label{eq:20}
\end{equation}
were used as the problem criteria; the expressions 
\begin{equation}
a_{ij}(y_1, y_2) = \sin(\pi i y_1) \sin(\pi j y_2), b_{ij}(y_1, y_2) = \cos(\pi i y_1) \cos(\pi j y_2)
\end{equation}
were defined in the range $0 \leq y_1,y_2 \leq 1$, the parameters $1 \leq A_{ij},B_{ij},C_{ij},D_{ij} \leq 1$ were the independent random numbers. The minimization of such functions arises, for example, in the problem of evaluation of the maximum strain (determining the strength) in a thin plate under the normal load.

In Fig. 1, the contour plots of three functions of this family are shown. As one can see, the functions of such kind are essentially multiextremal. 
\begin{figure}[t]
\includegraphics[width=\textwidth]{fig1}
\caption{Contour plots of three multiextremal functions from the test optimization problem family} 
\label{fig:1}
\end{figure}

In \cite{c9}, the proposed approach using the min-max scalarization scheme was compared with the well-known multicriteria optimization methods:
\begin{itemize}
	\item The Monte-Carlo (MC) method, where the trial points are selected within the search domain $D$ randomly and uniformly,
	\item The genetic algorithm SEMO from the PISA library,
	\item The non-uniform coverage (NUC) method,
	\item The bi-objective Lipschitz optimization (BLO) method.
\end{itemize}

In this paper, the efficiency of the proposed approach is compared with other scalarization schemes: the method of successive concessions and the reference point method.

In order to draw more justified conclusions on the efficiency of the developed approach, we solved 100 multicriteria problems formed using the multiextremal functions of the family (\ref{eq:20}). To solve each problem, a total of 50 different coefficients $\alpha$ of the convolutions (\ref{eq:4})-(\ref{eq:6}) were used. The values of the coefficient $\alpha$ are uniformly distributed within the interval $[0,1]$. The parameter values in our experiments were taken as follows: the reliability parameter $r=2,3$, the required accuracy $\varepsilon = 0,01$. All the results presented below were averaged over the family of the  problems solved.

In the first experiment, multicriteria problems were solved using all the considered scalarization schemes (\ref{eq:4})-(\ref{eq:6}), for which a set of different scalarization coefficients was taken. The calculations were performed without using the accumulated search information $A_k$ from (\ref{eq:12}). The numerical results are shown in Table \ref{tab:1}, where the columns labeled ''K'' contain the average number of iterations; in the columns labeled as ''S'', the achieved speedup of the parallel computations is shown. As follows from the presented results of our experiments, the parallel algorithm MPGSA based on the MGGSA method \cite{c10,c16} showed a high efficiency. When employing 16 cores, the speedup was more than 12.5 times for any type of convolutions (\ref{eq:4})-(\ref{eq:6}).

\begin{table}[t]
\caption{Average number of iterations and the speedup of parallel computations when solving a single MCO problem without reusing the search information}\label{tab:1}
\centering
\begin{tabular}{clcclcclcclcclcc}
\hline
           &  & \multicolumn{14}{c}{Number of cores}                                                                                               \\ \cline{3-16}
Convolution&  & \multicolumn{2}{c}{1} &  & \multicolumn{2}{c}{2} &  & \multicolumn{2}{c}{4} &  & \multicolumn{2}{c}{8} &  & \multicolumn{2}{c}{16} \\ \cline{3-4} \cline{6-7} \cline{9-10} \cline{12-13} \cline{15-16} 
           &  & K            & S      &  & K          & S        &  & K          & S        &  & K          & S        &  & K          & S         \\ \cline{1-1} \cline{3-4} \cline{6-7} \cline{9-10} \cline{12-13} \cline{15-16} 
$F_1 (\lambda,y)$ from (\ref{eq:4}) &  & 13 221       & 1      &  & 7 187      & 1,8      &  & 3 456      & 3,8      &  & 1 897      & 7,0      &  & 1 058      & 12,5      \\
$F_2 (\delta,y)$ from (\ref{eq:5})  &  & 16 002       & 1      &  & 9 458      & 1,7      &  & 4 113      & 3,9      &  & 2 356      & 6,8      &  & 1 204      & 13,3      \\
$F_3 (\theta,y)$ from (\ref{eq:6})  &  & 14 041       & 1      &  & 7 866      & 1,8      &  & 3 558      & 3,9      &  & 1 982      & 7,1      &  & 1 068      & 13,2      \\ \hline
\end{tabular}
\end{table}

\begin{table}[t]
\caption{Average number  of iterations and speedup of parallel computations when solving a single MCO problem with the reuse of the search information}\label{tab:2}
\centering
\begin{tabular}{clcclcclcclcclcc}
\hline
           &  & \multicolumn{14}{c}{Number of cores}                                                                                               \\ \cline{3-16}
Convolution&  & \multicolumn{2}{c}{1} &  & \multicolumn{2}{c}{2} &  & \multicolumn{2}{c}{4} &  & \multicolumn{2}{c}{8} &  & \multicolumn{2}{c}{16} \\ \cline{3-4} \cline{6-7} \cline{9-10} \cline{12-13} \cline{15-16} 
           &  & K            & S      &  & K          & S        &  & K          & S        &  & K          & S        &  & K          & S         \\ \cline{1-1} \cline{3-4} \cline{6-7} \cline{9-10} \cline{12-13} \cline{15-16} 
$F_1 (\lambda,y)$ from (\ref{eq:4}) & & 1 193 & 1 & & 595  & 2,0  & & 300 & 4,0  & & 173 & 6,9 & & 103 & 11,5 \\
$F_2 (\delta,y)$ from (\ref{eq:5})  & & 2 021 & 1 & & 934  & 2,2  & & 475 & 4,3  & & 272 & 7,4 & & 147 & 13,7 \\
$F_3 (\theta,y)$ from (\ref{eq:6})  & & 990	  & 1 & & 491  & 2,0  & & 249 & 4,0  & & 145 & 6,8 & & 90	 & 10,9 \\ \hline
\end{tabular}
\end{table}

The results presented in Table \ref{tab:2} show  that by reusing the search information the amount of performed computations can be reduced at least 10-fold.

The efficiency of the developed approach becomes more evident when the obtained reduction of executed optimization iterations is shown in comparison with the initial sequential algorithm, which does not use the search information (Table \ref{tab:3}). As follows from the results of performed experiments, the overall reduction when using 16 computation cores always exceeded 100.

\begin{table}[t]
\caption{Overall reduction of executed iterations provided by using the developed approach for solving a MCO problem}\label{tab:3}
\centering
\begin{tabular}{cccccc}
\hline
                  & \multicolumn{5}{c}{Number of cores} \\ \cline{2-6} 
Convolution       & 1     & 2     & 4    & 8    & 16    \\ \hline
$F_1 (\lambda,y)$ from (\ref{eq:4}) & 11,1  & 22,2  & 44,1 & 76,2 & 127,5 \\
$F_2 (\delta,y)$  from (\ref{eq:5}) & 7,9   & 17,1  & 33,7 & 58,7 & 108,2 \\
$F_3 (\theta,y)$  from (\ref{eq:6}) & 14,2  & 28,6  & 56,3 & 96,3 & 154,9 \\ \hline
\end{tabular}
\end{table}

\section{Conclusion}
In the present paper, a new approach to parallel computations for solving time-consuming multicriteria global optimization problems is presented. This approach is based on various methods for scalarization of vector criteria, on dimensionality reduction with the use of the Peano space-filling curves, and on efficient global search algorithms.

A novel contribution consists in the development of an approach permitting to use different criteria scalarization schemes with the account of different optimality requirements to desirable decisions. For the scalarization schemes considered, a general method of parallel computations was proposed and the efficiency of the developed approach was evaluated computationally. Our numerical experiments confirmed the efficiency of the proposed approach -- for example, the speedup of parallel computations when using 16 computational cores always exceeded 100.

In further investigations, we intend to perform numerical experiments on parallel solving the multicriteria optimization problems for a larger number of efficiency criteria and for larger dimensionality. Parallel computations on computational nodes with distributed memory can be considered as well.

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}

\bibitem{c1} 	Marler, R.T., Arora, J.S.: Multi-Objective Optimization: Concepts and Methods for Engineering. In: VDM Verlag (2009).
\bibitem{c2} 	Ehrgott, M.: Multicriteria Optimization. In: Springer (2nd ed., 2010). 
\bibitem{c3} 	Collette, Y., Siarry, P.: Multiobjective Optimization: Principles and Case Studies (Decision Engineering). In: Springer (2011).
\bibitem{c4} 	Pardalos, P.M., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-Convex Multi-Objective Optimization. In: Springer (2017). 
\bibitem{c5} 	Modorskii, V.Y., Gaynutdinova, D.F., Gergel, V.P., Barkalov, K.A.: Optimization in design of scientific products for purposes of cavitation problems. In: AIP Conference Proceedings, \textbf{1738}, pp. 400013 (2016). \doi{10.1063/1.4952201}
\bibitem{c6} 	Strongin, R.G., Gergel, V.P.: Parallel computing for globally optimal decision making. In: Lecture Notes in Computer Science, \textbf{2763}, pp. 76--88 (2003).
\bibitem{c7} 	Gergel, V., Kozinov, E.: Efficient methods of multicriterial optimization based on the intensive use of search information. In: Springer Proceedings in Mathematics and Statistics, \textbf{197}, pp. 27--45 (2017). \doi{10.1007/978-3-319-56829-4\_3}
\bibitem{c8} 	Barkalov, K., Gergel, V., Lebedev, I.: Solving global optimization problems on GPU cluster. In: AIP Conference Proceedings, \textbf{1738}, pp. 400006 (2016).
\bibitem{c9} 	Gergel, V.P., Kozinov, E.A.: Efficient multicriterial optimization based on intensive reuse of search information. In: J Glob Optim., \textbf{71}(1), pp. 73--90, (2018).
\bibitem{c10}	Strongin, R., Sergeyev, Ya.: Global optimization with non-convex constraints. Sequential and parallel algorithms. In: Kluwer Academic Publishers, Dordrecht (2nd ed. 2013, 3rd ed. 2014).
\bibitem{c11}	Sergeyev Y.D., Strongin R.G., Lera D.: Introduction to global optimization exploiting space-filling curves. In: Springer (2013).
\bibitem{c12}	Zhigljavsky, A., {\v Z}ilinskas, A.: Stochastic Global Optimization. In: Springer (2008).
\bibitem{c13}	Locatelli, M., Schoen, F.: Global Optimization: Theory, Algorithms, and Applications. In: SIAM (2013).
\bibitem{c14}	Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. In: Princeton University Press (2016).
\bibitem{c15}	Gergel, V.P., Sidorov, S.V.: A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems. In: Parallel Computing Technologies / LNCS, \textbf{9251}, pp. 505--515 (2015).
\bibitem{c16}	Gergel, V., Kozinov, E.: Parallel computing for time-consuming multicriterial optimization problems. In: LNCS, \textbf{10421}, pp. 446--458 (2017). 
\bibitem{c17}	Sergeyev Y.D., Kvasov D.E.: A deterministic global optimization using smooth diagonal auxiliary functions. In: Communications in Nonlinear Science and Numerical Simulation, \textbf{21}(1-3), pp. 99--111 (2015).
\bibitem{c18}	Gergel, V., Grishagin, V., Gergel, A.: Adaptive nested optimization scheme for multidimensional global search. In: JOGO, \textbf{66}(1), pp. 35--51 (2016).

\end{thebibliography}
\end{document}

### 2019-11-19 2019/RSCD_Gergel_Kozinov/RSCD_2019.tex
Добавил исходники статей по ГО за 2019г. (Гергель Козинов)

--- 2018-12-11 2019/ICCS_MCO/ICCS_2019.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% This is a general template file for the LaTeX package SVJour3
-% for Springer journals.          Springer Heidelberg 2010/09/16
-%
-% Copy it to a new file with a new name and use it as the basis
-% for your article. Delete % signs as needed.
-%
-% This template includes a few options for different layouts and
-% content for various journals. Please consult a previous issue of
-% your journal as needed.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% First comes an example EPS file -- just ignore it and
-% proceed on the \documentclass line
-% your LaTeX will extract the file if required
-%\begin{filecontents*}{example.eps}
-%%!PS-Adobe-3.0 EPSF-3.0
-%%%BoundingBox: 19 19 221 221
-%%%CreationDate: Mon Sep 29 1997
-%%%Creator: programmed by hand (JK)
-%%%EndComments
-%gsave
-%newpath
-%  20 20 moveto
-%  20 220 lineto
-%  220 220 lineto
-%  220 20 lineto
-%closepath
-%2 setlinewidth
-%gsave
-%  .4 setgray fill
-%grestore
-%stroke
-%grestore
-%\end{filecontents*}
-%
-\RequirePackage{fix-cm}
-%
-%\documentclass{svjour3}                     % onecolumn (standard format)
-%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
 \documentclass[runningheads]{llncs}
-%\documentclass[twocolumn]{svjour3}          % twocolumn
 %
@@
 \usepackage{mathtools,amssymb}
+\usepackage{bbding}
 %
@@
 
-\title{A Highly Parallel Approach for Solving Computationally Expensive Multicriteria Optimization Problems
-\thanks{This work has been supported by Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using superior-performance supercomputers.''}
-}
+\title{A Highly Parallel Approach for Solving Computationally Expensive Multicriteria Optimization Problems}
 
@@
 
-\author{Victor Gergel\orcidID{0000-0002-4013-2329} \and Evgeny Kozinov \orcidID{0000-0001-6776-0096}}
+\author{Victor Gergel\orcidID{0000-0002-4013-2329} (\Envelope) \and Evgeny Kozinov \orcidID{0000-0001-6776-0096}}
 
@@
 \min{(\phi(x) = F(\lambda, y(x)), y(x) \in D, x \in [0,1])},\\
-F(\lambda,y(x))=\max⁡{(\lambda_i f_i(y(x)), 1 \leq i \leq s)},\\
+F(\lambda,y(x))=\max{(\lambda_i f_i(y(x)), 1 \leq i \leq s)},\\
 \lambda=(\lambda_1,\lambda_2,\dots,\lambda_s) \in \Lambda \subset R^s : \sum_{i=1}^s{\lambda_i}=1, \lambda_i \geq 0, 1 \leq i \leq s
@@
 
-Multiextremal optimization is a research area being actively developed -- its state of the art is presented, for example, in \cite{c7,c9,c10}.
+Multiextremal optimization is a research area being actively developed -- its state of the art is presented, for example, in \cite{c7,c9,c10,c60}.
 The information-statistical theory of global optimization is considered to be one of the promising approaches \cite{c7}.
@@
 
-Within the proposed approach, for solving the reduced one-dimensional multiextremal optimization subproblems (\ref{eq:3}) it is proposed to use the well-known Parallel Algorithm of Global Search (PAGS) developed within the framework of the information-statistical theory of the multiextremal optimization \cite{c7}.
+Within the proposed approach, for solving the reduced one-dimensional multiextremal optimization subproblems (\ref{eq:3}) it is proposed to use the well-known Parallel Algorithm of Global Search (PAGS) developed within the framework of the information-statistical theory of the multiextremal optimization \cite{c7,c25}.
 This method has a good theoretical background and has demonstrated its efficiency as compared to other global search algorithms (see also the results of numerical experiments in Section \ref{sec:4}).
@@
 
-Rule 5. Perform new trials at the points $x^{k+j}$, $1 \leq j \leq p$ placed in the intervals with the maximum characteristics from (\ref{eq:11}) according to the expressions
+\textit{Rule 5.} Perform new trials at the points $x^{k+j}$, $1 \leq j \leq p$ placed in the intervals with the maximum characteristics from (\ref{eq:11}) according to the expressions
 
@@
 \begin{eqnarray} \label{eq:14}
-z_k^*=\min⁡{\{z_i:1\leq i \leq k\}}.
+z_k^*=\min{\{z_i:1\leq i \leq k\}}.
 \end{eqnarray}
@@
 \begin{eqnarray} \label{eq:15}
-z_i'=\max{⁡(\lambda_j' f_i^j,1\leq j \leq s)},1 \leq i \leq k.
+z_i'=\max{(\lambda_j' f_i^j,1\leq j \leq s)},1 \leq i \leq k.
 \end{eqnarray}
@@
 \begin{equation*}
-F(\lambda,y(x^{k+j})) \leq F(\lambda,y^* )+\varepsilon,1 \leq j \leq p,  F(\lambda,y^*) = \min_{y\in D_\delta}⁡{F(\lambda,y)}.
+F(\lambda,y(x^{k+j})) \leq F(\lambda,y^* )+\varepsilon,1 \leq j \leq p,  F(\lambda,y^*) = \min_{y\in D_\delta}{F(\lambda,y)}.
 \end{equation*}
@@
 \begin{equation}\label{eq:18}
-\Delta=\frac{1}{M} \sum_{y\in PD_\delta(f,D)}{d(y)}, d(y)=\min_{\overline{y}\in PD_{\Omega_k}(f,D)}{⁡\|y-\overline{y}\|},
+\Delta=\frac{1}{M} \sum_{y\in PD_\delta(f,D)}{d(y)}, d(y)=\min_{\overline{y}\in PD_{\Omega_k}(f,D)}{\|y-\overline{y}\|},
 \end{equation}
@@
 
+\section*{Acknowledgements}
+This work has been supported by Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using superior-performance supercomputers.''
+
 %\begin{acknowledgements}
@@
 505--515 (2015)
-\bibitem{c15}	Gergel, V.: An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems.
-In: 2nd International Conference on Mathematics and Com-puters in Sciences and in Industry, pp.
-13--18 (2015)
+\bibitem{c15}	Barkalov, K., Gergel, V., Lebedev, I.: Solving global optimization problems on GPU cluster. AIP Conference Proceedings, \textbf{1738}, pp. 400006 (2016) \doi{10.1063/1.4952194}
 \bibitem{c16}	Gergel, V., Kozinov, E.: Efficient Multicriteria Optimization Based on Intensive Reuse of Search Information. In: Journal of Global Optimization, \textbf{71}(1), pp. 73--90 (2018) 
@@
 \bibitem{c50}	Borisenko, A., Gorlatch, S.: Parallelizing Metaheuristics for Optimal Design of Multiproduct Batch Plants on GPU. In: LNCS, Springer, Cham, \textbf{10421}, pp. 405--417 (2017)
+\bibitem{c25}	Strongin, R.G., Gergel, V.P.: Parallel computing for globally optimal decision making. In: Lecture Notes in Computer Science, \textbf{2763}, pp. 76--88 (2003)  
+\bibitem{c60}	Sakharov, M.K., Karpenko, A.P.:Adaptive Load Balancing in the Modified Mind Evolutionary Computation Algorithm In: Supercomputing Frontiers and Innovations, \textbf{5}(4), pp.
+5--14 (2018)
 

### 2019-11-19 2019/WCGO_Gergel_Kozinov/MCO_WCGO_2019_03_29_latex.tex
Добавил исходники статей по ГО за 2019г. (Гергель Козинов)

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Multistage Global Search Using Various Scalarization Schemes in Multicriteria Optimization Problems\thanks{This research was supported by the Russian Science Foundation, project No 16-11-10150 ''Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance.''}}
%
\titlerunning{Multistage Global Search in Multicriteria Optimization Problems}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Victor Gergel\orcidID{0000-0002-4013-2329} \and
Evgeniy Kozinov\orcidID{0000-0001-6776-0096}}
%
\authorrunning{V. Gergel et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
\email{gergel@unn.ru} \email{evgeny.kozinov@itmm.unn.ru}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper, an approach, in which the decision making problems are reduced to solving the multicriteria time-consuming global optimization problems is proposed. The developed approach includes various methods of scalarization of the vector criteria, the dimensionality reduction with the use of the Peano space-filling curves and the efficient global search algorithms. In the course of computations, the optimization problem statements and the applied methods of the criteria scalarization can be altered in order to achieve more complete compliance to available requirements to the optimality. The overcoming of the computational complexity of the developed approach is provided by means of the reuse of the whole search information obtained in the course of computations. The performed numerical experiments have confirmed the reuse of the search information to allow reducing essentially the amount of computations for solving the global optimization problems.

\keywords{Decision making \and Multicriteria optimization \and Criteria scalarization \and Global optimization with nonlinear constraints \and Numerical experiment.}
\end{abstract}
%
%
%
\section{Introduction}	
\label{sec:01}
The multicriteria optimization (MCO) problems, which are used as the statements of the decision making problems often are the field of extensive research -- see, for example, the monographs \cite{c1,c2,c3} and the reviews of the scientific and applied results in this area \cite{c4,c5}.

Usually, the finding of the efficient (non-dominated) decisions, in which the improvement of the values with respect to any criteria cannot be achieved without the worsening of the values with respect to other criteria is understood as the solution of a MCO problem. In the most general case, when solving the MCO problems, it can appear to be necessary to obtain a complete set of the efficient decisions (the Pareto set). However, the finding of all efficient decisions may require a considerable amount of computations, and the set of obtained decisions may appear to be quite large. As a result, the approaches, which obtain the more limited set of efficient decisions are applied wider. Among such approaches, there are various kinds of the criteria convolutions, the lexicographic optimization methods, the algorithms of searching the best approximation to the given prototypes, etc. All methods mentioned above allow accounting for the specific features of the MCO problems being solved and satisfy the requirements to the optimality from the decision maker.

The present work is devoted to the solving of the MCO problems, which are used for the description of the complex decision making problems, in which the criteria of efficiency may have a multiextremal form, and the determining of the values of the criteria and constraints may require a large amount of computations. Also, it was assumed that in the course of computations it is possible to change the problem statement, the methods, and the parameters of solving the MCO problem that results in the necessity of the multiple solving of the global optimization problems.

The practical use of this approach implies the overcoming of a considerable computational complexity of the decision making problems that can be provided by means of the use of highly efficient global optimization methods and the complete utilization of the search information obtained in the course of computations.

In the present paper, the results of investigations on the generalization of the decision making problem statements \cite{c6,c7} and on the development of the highly efficient global optimization methods utilizing the whole search information obtained in the course of computations \cite{c8,c9,c10} are presented.

Further structure of the paper is as follows. In Section \ref{sec:02}, the statement of the decision making problems based on multistage multicriteria global optimization are presented. In Section \ref{sec:03}, a general scheme of the criteria scalarization is proposed. In Section \ref{sec:04}, the search information obtained in the course of computations is considered. In Section \ref{sec:05}, an efficient algorithm for solving the time-consuming global optimization problems with the nonlinear constraints is presented. Section \ref{sec:06} presents the results of numerical experiments confirming the developed approach to be promising. In Conclusion, the obtained results are discussed and possible main directions of further research are outlined.

\section{Multistage multicriteria optimization problem statement}	
\label{sec:02}
For the formal description of the process of solving the complex decision making problems, the following generalized two-phase scheme is proposed.

1.	In the most general form, a decision making problem is defined by means of a vector function of characteristics 
\begin{equation}\label{eq:1}
	w(y)=(w_1(y), w_2(y), \dots, w_M (y)), y \in D
\end{equation}
where $y=(y_1,y_2,\dots, y_N )$ is the vector of varied parameters and $D \subset R^N$ is the search domain, which is usually an $N$-dimensional hypercube
\begin{equation}\label{eq:2}
	D=\{y \in R^N:a_i \leq y_i \leq b_i,1 \leq i \leq N\}
\end{equation}
for given vectors $a$ and $b$.

It is supposed, that the values of characteristics $w(y)$ are non-negative, and the decreasing of these ones corresponds to the increasing of the efficiency of the chosen decisions. Also, it is supposed that the characteristics $w_j (y)$, $1 \leq j \leq M$, may be multiextremal, and the determining of their values may require large enough amount of computations. Besides, the characteristics $w_j (y)$, $1 \leq j \leq M$, are supposed to satisfy the Lipschitz condition
\begin{equation}\label{eq:3}
	|w_j (y_1 )-w_j (y_2 )| \leq L_j \|y_1-y_2 \|, 1 \leq j \leq M
\end{equation}
where $L_j$ is the Lipschitz constant for the characteristic $w_j(y)$ , $1 \leq j \leq M$, and $\|*\|$ denotes the Euclidean norm in $R^N$.

2.	Then, a MCO problem is formulated on the basis of the statement considered above. For this purpose, a vector criterion of efficiency is selected among the characteristics $w_j (y)$, $1 \leq j \leq M$, from (\ref{eq:1}) 
\begin{equation}\label{eq:4}
	f(y)=(f_1 (y),f_2 (y),\dots,f_s (y)), f_j (y)=w_{i_j} (y), 1\leq j \leq s, 1\leq i_j \leq M
\end{equation}
and the vector function of constraints
\begin{equation}\label{eq:5}
	g(y)=(g_1 (y),g_2 (y),\dots,g_m (y)), g_l (y)=w_{i_l} (y)-q_l, 1\leq l\leq m, 1 \leq i_l \leq M
\end{equation}
where $q_l>0$, $1 \leq l \leq m$, are the allowances on the feasible values of characteristics $w_j (y)$, $1 \leq j \leq M$.

The formulated criteria of efficiency and constraints allow to define a multicriteria optimization problem
\begin{equation}\label{eq:6}
	P : f(y)\to min, y \in Q
\end{equation}
where Q is the feasible search domain
\begin{equation}\label{eq:7}
	Q=\{ y\in D : g(y) \leq 0 \}.
\end{equation}

In development of this scheme of the MCO problem statement, further an opportunity of simultaneous formulation of several MCO problems 
\begin{equation}\label{eq:8}
	\mathbb{P}_t=\{ P_1,P_1,\dots,P_t \},
\end{equation}
will be allowed, the set of which can be varied in the course of computations by means of adding new or by removing already existing optimization problems.

In general, the proposed scheme of the optimal decision search process (\ref{eq:1})-(\ref{eq:8}) defines a new class of the optimization problems -- the multistage multicriteria global optimization (MMGO) problems.

\section{Reduction of the multistage multicriteria search to the scalar one-dimensional global optimization problems}	
\label{sec:03}
One of the widely used approaches to solving the MCO problems is to transform the vector criterion into some general scalar criterion of efficiency that allows using a large set of already existing optimization methods for solving the global optimization problems. Among the possible scalarization methods, there are, for example, the weighted sum method, the compromise programming method, the weighted minimax method, and many other methods -- see, for example, \cite{c2,c3}.

In the general form, the statement of the global optimization problems generated as a result the MCO problem criteria scalarization can be represented as follows:
\begin{equation}\label{eq:9}
	\min{\varphi(y) = F(\alpha,y)}, g(y) \leq 0, y \in D,
\end{equation}
where $F$ is the scalar objective function, $\alpha$ is the vector of parameters of the applied criteria scalarization method, $g(y)$ are the constraints of the MCO problem from (\ref{eq:6}), and $D$ is the search domain from (\ref{eq:2}).

Particular forms of the function $F(\alpha,y)$ are defined by the criteria scalarization methods applied. For example, the following scalarization methods are possible.

1.	In the case of equal importance of the criteria $f_i$, $1 \leq i \leq s$, the minimax convolution scheme (MMC) \cite{c3} can be applied:
\begin{equation}\label{eq:10}
\begin{split}
  F^1 (\lambda,y)=\max{⁡(\lambda_i f_i(y), 1 \leq i \leq s)},  \\
  \lambda=(\lambda_1,\lambda_2,\dots,\lambda_s) \in \Lambda \subset R^s:\sum_{i=1}^s{\lambda_i =1},\lambda_i \geq 0, 1 \leq i \leq s.
\end{split}
\end{equation}

2.	In the case of arrangement of the criteria in importance order, the method of successive concessions (MSC) \cite{c2,c3} can be used. In this case, the scalar objective function F can be set as follows
\begin{equation}\label{eq:11}
	\min F^2 (\lambda,y)=f_s (y),f_i (y) \leq f_i^{min}+\delta_i (f_i^{max}-f_i^{min} ),1 \leq i<s,g(y)\leq 0,y\in D,
\end{equation}
where $f_i^{min}$ and $f_i^{max}$ are the minimum and maximum values of the criteria $f_i (y)$, $1\leq i<s$, in the domain $D$ respectively, and $0 \leq \delta_i \leq 1$, $1\leq i<s$, are the concessions with respect to each criterion. As before, the values of concessions $0 \leq \delta_i\leq 1$, $1 \leq i<s$, can be varied in the course of computations. The quantities $f_i^{min}$ and $f_i^{max}$,  $1 \leq i<s$, the values of which may be unknown \textit{a priori}, can be replaced by the minimum and maximum estimates of the criteria values computed using the available search information.

3.	In the case of availability of any estimates of the criteria values of the required decision (for example, based on an ideal decision or on any existing prototype) the MCO problem solution may consist in finding an efficient decision the most completely matching given values of optimality (the reference point method, RPM). Such a problem can be formulated in the form of a scalar optimization problem
\begin{equation}\label{eq:12}
	\min{F^3 (\lambda,y) = 1 / s \sum_{i=1}^s{\theta_i (f_i (y)-f_i^* )^2 } },g(y)\leq 0,y \in D	
\end{equation}
where the objective function $F^3 (\lambda,y)$ is the standard deviation of the decision $y \in D$ from the sought ideal decision, and the quantities $0 \leq \theta_i \leq 1$, $1 \leq i<s$, are the magnitudes of importance of approximations with respect to each particular variable $y_i$, $1\leq i\leq N$.

Within the framework of the developed approach, it is possible to change the used scalarization methods (\ref{eq:10})-(\ref{eq:12}) and/or altering the parameters of convolutions $\lambda$, $\delta$ and $\theta$. Such variations expand the set of the MCO problems $\mathbb{P}$ from (\ref{eq:8}) necessary for solving the initial decision making problem into a wider set of the scalar global optimization problems (\ref{eq:9})
\begin{equation}\label{eq:13}
	\mathbb{F}_T=\{ F_i (\alpha_i,y): 1\leq i \leq T \},
\end{equation}
in which each problem $P\in \mathbb{P}$ from (\ref{eq:8}) can correspond to several global optimization problems (\ref{eq:9}).

In the developed approach, one more step of converting the problems being solved $F(\lambda,y)$ from (\ref{eq:9}) is applied, namely the dimensionality reduction is performed with the use of the Peano space-filling curves (evolvents) $y(x)$ providing an unambiguous mapping of the interval [0,1] onto an $N$-dimensional hypercube $D$ \cite{c11,c12}. As a result of such reduction, the multidimensional global optimization problem (\ref{eq:9}) is reduced to a one-dimensional problem
\begin{equation}\label{eq:14}
	F(\alpha,y(x^* ))=\min{\{ F(\alpha,y(x)) : g(y(x))\leq 0, x\in[0,1] \}.}
\end{equation}

The dimensionality reduction allows applying many well known highly efficient one-dimensional global optimization algorithms for solving the problems (\ref{eq:9}) (after performing necessary generalization) -- see, for example, \cite{c11,c12,c13,c14,c15}.

\section{Computational complexity reduction of the multistage multicriteria search on the basis of the reuse of the search information}	
\label{sec:04}

The numerical solving of the global optimization problems (\ref{eq:9}) is usually implemented by the successive computing the values of characteristics $w(y)$ at the points $y^i$, $0\leq i\leq k$, of the search domain $D$ \cite{c11,c14}. The data obtained as a result of computations can be represented in the form of the matrix of the search information
\begin{equation}\label{eq:15}
	\Omega_k=\{ (y^i,w^i=w(y^i ))^T: 1\leq i \leq k \}.
\end{equation}

As a result of scalarization (\ref{eq:9}) and the use of the dimensionality reduction (\ref{eq:14}), the set $\Omega_k$ from (\ref{eq:15}) can be transformed into the form of the matrix of the search state
\begin{equation}\label{eq:16}
	A_k={ (x_i,z_i,g_i,l_i )^T: 1\leq i \leq k },
\end{equation}
where $x_i$, $1\leq i\leq k$, are the reduced points of the executed global search iterations, $z_i$, $g_i$, $1\leq i\leq k$, are the values of the scalar criterion and constraints of current optimization problem $F(\alpha,y(x))$ at these points, and $l_i$, $1\leq i\leq k$, are the indices of global search iterations, in which the points $x_i$, $1\leq i\leq k$, were computed.

The availability of the set $\Omega_k$ from (\ref{eq:15}) allows reducing the results of all preceding computations $z_i$, $1\leq i\leq k$ from the matrix $A_k$ to the values of the next optimization problem $F(\alpha,y(x))$ from (\ref{eq:9}) without any repeated time-consuming computations of the values of $w(y)$ from (\ref{eq:1}), i. e.
\begin{equation}\label{eq:17}
	w_i \xrightarrow{\alpha,P} (z_i,g_i ), 1\leq i\leq k, \forall \alpha, P\in \mathbb{P}
\end{equation}

The reuse of the search information can provide a gradual decreasing of the amount of computations when solving every next optimization problem down to the execution of few iterations only to find the next efficient decision.

\section{Efficient solving the multistage multicriteria optimization problems with nonlinear constraints}	
\label{sec:05}
Within the framework of developed approach, the efficient method was applied for solving the global optimization problems with the nonlinear constraints \cite{c11}. The essence of the approach is the constructing of a problem with some aggregated unconstrained objective function, the solving of which leads to the solution of the initial problem (\ref{eq:9}) -- more detailed description of the approach will be given below.

Let us introduced a simpler notation for the reduced problems (\ref{eq:14}) as
\begin{equation}\label{eq:18}
%\begin{split}
	\min{\{g_{m+1} (x):g_i (y(x)) \leq 0, 1\leq i \leq m,x \in [0,1]\} }, g_{m+1} (x)=F(\lambda,y(x)).
%\end{split}
\end{equation}

The problem (\ref{eq:18}) can be considered in the statement of partial computability when each function $g_j (y)$, $1 \leq j \leq m+1$ is defined and computable in certain subinterval $\Delta_j \subset [0,1]$ only where
\begin{equation}\label{eq:19}
	\Delta_1=[0,1], \Delta_{j+1}=\{ x \in \Delta_j  : g_j (y(x))\leq 0 \}, 1\leq j\leq m.
\end{equation}

Taking into account the conditions (\ref{eq:19}), the objective function of the problem (\ref{eq:18}) can be represented in the form 
\begin{equation}\label{eq:20}
	\varphi(x^*)=\min{\{ g_{m+1} (y(x)) : x\in \Delta_{m+1} \}},
\end{equation}
on the basis of which the problem (\ref{eq:18}) can be transform to the problem of minimizing the aggregated function 
\begin{equation}\label{eq:21}
\begin{split}
	\min{\{ \Phi(y(x))=g_v(y(x)), v=v(x), x\in[0,1] \}}, \\	
	1\leq v=v(x)\leq m+1, g_v (y(x))>0, g_j (y(x))\leq 0, 1\leq j\leq v-1.
\end{split}
\end{equation}

The index $v=v(x)$, $1\leq v \leq m+1$ defines the first violated constraint at the point $x$ in successive calculation of constraints.

Within the framework of the developed approach, the algorithm of global constrained optimization (AGCO), which is considered in details in \cite{c11} is applied for solving the global unconstrained problems (\ref{eq:21}).

\section{Results of numerical experiments}	
\label{sec:06}
The numerical experiments have been carried out using the computational nodes of Lobachevsky supercomputer at Nizhni Novgorod State University. The peak performance of the supercomputer was 573 Tflops, each computational node was equipped with Intel Sandy Bridge E5-2660 processor 2.2 GHz, 64 Gb RAM.

Within the framework of the present paper, the construction of a numerical approximation (PDA) of the Pareto domain (PD) was understood as a solution of a MCO problem. In order to evaluate the efficiency of constructing the PDA, two main indicators applied widely were used: the completeness of coverage of the Pareto domain (hypervolume index, HV) and the uniformity of distribution of the numerical estimates of the efficient decisions (distribution uniformity index, DU) \cite{c10}. The higher values of the index HV and the lower values of the index DU corresponds to the better approximation.

Within the executed experiments, the two-dimensional bi-criterial MCO problems were used, the criteria of which were defined with the use of the family of multiextremal functions \cite{c11}. The functions of this family are defined as follows
\begin{equation}\label{eq:22}
\begin{split}
f(y_1,y_2 )= -(AB+AC)^{1/2},\\
AB=(\sum_{i=1}^7 {\sum_{j=1}^7{[A_{ij} a_{ij} (y_1,y_2 )+B_{ij} b_{ij} (y_1,y_2 )]}} )^2,\\
CD=(\sum_{i=1}^7 {\sum_{j=1}^7{[C_{ij} a_{ij} (y_1,y_2 )-D_{ij} b_{ij} (y_1,y_2 )]}} )^2,
\end{split}
\end{equation}
where the expressions 
\begin{equation*}
	a_{ij} (y_1, y_2)=sin(\pi iy_1 )  sin(\pi jy_2 ), b_{ij} (y_1,y_2 )=cos(\pi iy_1 ) cos(\pi jy_2 )
\end{equation*}
are defined in the area $0\leq y_1,y_2\leq 1$, and the parameters $-1 \leq A_{ij},B_{ij},C_{ij},D_{ij} \leq 1$ are the independent random numbers distributed uniformly in the same interval.

The initial series of the numerical experiments was performed to evaluate the positive effect (the reducing of the number of the executed iterations) due to the reuse of the search information obtained in the course of computations (see Section \ref{sec:04}). When conducting the experiments, the AGCO algorithm was applied sequentially with all scalarization methods considered in Section Section \ref{sec:03}: the minimax convolution of the criteria (\ref{eq:10}), the method of successive concessions (\ref{eq:11}), and the reference point method (\ref{eq:12}). For each method mentioned above, 50 subproblems with various values of the parameters $\lambda$, $\delta$ and $\theta$ correspondingly have been solved. In order to evaluate the indicator HV of the completeness of the PDA approximation, the reference point (-4,-4) was used. For the AGCO algorithms, the reliability parameter $r=2.3$ was used; the accuracy in the stopping condition was set as $\varepsilon=0.01$.

The results of performed experiments are presented in Table 1. The column ''Iters'' contains the total number of iterations (the number of computations of the criteria values) required for solving the problem (\ref{eq:22}). The column ''PDA'' shows the number of the obtained efficient decisions in the PDA. In the first part of the table (columns 2-5), the results of solving all subproblems without the use of the search information are shown. In the second part (columns 6-9) the results of solving the same subproblems but with the use of the whole search information obtained in the course of computations (i. e. when solving every next subproblem, the search information obtained when solving all preceding subproblems was utilized) are presented. In the column ''S'', the speedup obtained as a result of reducing the number of the global search iterations (the number of the computed criteria values) required to solve all 50 subproblems due to the reuse of the search information is shown.

The results of experiments presented in Table 1 demonstrate the reduction of the number of the executed iterations of the AGCO algorithm more than 9 times. At that, the indicator HV takes almost the same values independently on the usage of the search information. However, the indicator DU was better notably (except the method of successive concessions) at the computations without the use of the search information (similar remark can be expressed on the number of efficient decisions in the PDA as well) -- most likely, this effect is manifested due to essentially less number of the executed iterations when using the search information. One can note also that the reference point method has some advantage in the efficiency indicators (the number of points in PDA, indicators HV and DU). However, this method inferior the use of the minimax convolutions in the number of iterations required to solve all 50 subproblems essentially.

\begin{table}[t]
\centering
\caption{Efficiency of the reuse of the search information in solving the problem (\ref{eq:22}) using various criteria scalarization methods}
\label{tab:01}
\begin{tabular}{ccccclcccclc}
\hline
              & \multicolumn{9}{c}{Search information}                      &  &     \\ \cline{2-10}
Scalarization & \multicolumn{4}{c}{Not used} &  & \multicolumn{4}{c}{Used}  &  & S   \\ \cline{2-5} \cline{7-10}
method        & Iters  & PDA  & HV    & DU   &  & Iters & PDA & HV   & DU   &  &     \\ \cline{1-5} \cline{7-10} \cline{12-12} 
MMC (\ref{eq:10})      & 9124   & 152  & 53.4  & 0.64 &  & 922   & 44  & 52.7 & 0.91 &  & 9.9 \\
MSC (\ref{eq:11})      & 12802  & 88   & 52.8  & 1.16 &  & 1402  & 40  & 52.4 & 1.09 &  & 9.1 \\
RPM (\ref{eq:12})      & 12881  & 240  & 53.6  & 0.51 &  & 1323  & 56  & 53.0 & 0.83 &  & 9.7 \\ \hline
\end{tabular}
\end{table}

In the last experiment, solving a test MCO problem has been conducted with a use of the search information but the applied criteria scalarization methods varied in the course of computations. In this experiment, at the first stage of computations, the minimax criteria convolution was used in solving three subproblems (\ref{eq:10}) with the convolution coefficients (1,0), (0.5,0.5), and (0,1), correspondingly. At the second stage, the scalarization method was changed to the reference point method (12), where the estimate (-14,-7) was used as the reference point with the weighting coefficients (0.5,0.5). And finally, at the third stage, the method of successive concessions (\ref{eq:11}) was applied with the concession with respect to the first criterion $\delta=0.5$.

\begin{table}[ht]
\centering
\caption{Results of solving the problem (\ref{eq:22}) with altering the criteria scalarization methods in the course of computations}
\label{tab:02}
\begin{tabular}{lccc}
\hline
Stage of computations, criteria scalarization method & Total iters & New iters & PDA \\ \hline
1. MMC (10), three subproblems                       & 304         & 304       & 14  \\
2. RPM (12)                                          & 402         & 98        & 19  \\
3. MSC (11)                                          & 437         & 35        & 24  \\ \hline
\end{tabular}
\end{table}

The results of performed experiments are given in Table 2. The column ''Total iters'' contains the total number of executed global search iterations whereas the column ''New iters'' shows the points of iterations executed at particular stage of solving the problem only. As follows from the results presented in Table 2, the number of the global search iterations executed at separate stages of solving the MCO is continuously problem reduced (from 304 down to 35), and the reuse of the search information provides the computationally efficient opportunity for dynamic variations of the criteria scalarization methods applied in the course of the MCO problem solution.

The PDA approximations of Pareto set obtained at the sequentially executed stages are shown in Fig. 1.
\begin{figure}[ht]
	\centering
		\includegraphics[width=0.70\textwidth]{fig1}
	\caption{Approximations of the Pareto set PDA obtained at the sequentially executed stages of computations}
	\label{fig:1}
\end{figure}

\section{Conclusion}
An approach is proposed to transform the decision making problems to the multi-stage multicriteria time-consuming global optimization problems. As a key property of the proposed approach it is supposed that the optimization problem statements and the applied methods of the criteria scalarization can be changed in the course of computations. The computational complexity is reduced by means of the reuse of the computed search information. The performed numerical experiments have confirmed the developed approach is promising.

In further investigations it is intended to execute the numerical experiments on solving the problems for a larger number of efficiency criteria and for larger dimensionality. Parallel computations can be considered as well.

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}

\bibitem{c1}	Parnell, G.S., Driscoll, P.J., Henderson, D.L., editors.: Decision Making in Systems Engineering and Management. Wiley, New Jersey (2008).
\bibitem{c2}	Collette, Y., Siarry, P.: Multiobjective Optimization: Principles and Case Studies (Decision Engineering). Springer. (2011).
\bibitem{c3}	Pardalos, P.M., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-Convex Multi-Objective Optimization. Springer. (2017).
\bibitem{c4}	Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications. Surv. Math. Ind. \textbf{11}, 1--42 (2005).
\bibitem{c5}	Modorskii, V.Y., Gaynutdinova, D.F., Gergel, V.P., Barkalov, K.A.: Optimization in design of scientific products for purposes of cavitation problems. AIP Conference Proceedings, \textbf{1738}, 400013 (2016). \doi{10.1063/1.4952201}
\bibitem{c6}	Strongin, R.G., Gergel, V.P.: Parallel computing for globally optimal decision making. Lecture Notes in Computer Science, \textbf{2763}, 76--88 (2003).
\bibitem{c7}	Gergel, V.P., Kozinov, E.A.: Accelerating multicriterial optimization by the intensive exploitation of accumulated search data. AIP Conference Proceedings, \textbf{1776}, 090003 (2016). \doi{10.1063/1.4965367}
\bibitem{c8}	Gergel, V.: An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems. 2nd International Conference on Mathematics and Computers in Sciences and in Industry. (2015). \doi{10.1109/MCSI.2015.18}
\bibitem{c9}	Barkalov, K., Gergel, V., Lebedev, I.: Solving global optimization problems on GPU cluster. AIP Conference Proceedings, \textbf{1738}, 400006 (2016). \doi{10.1063/1.4952194}
\bibitem{c10}	Gergel, V.P., Kozinov, E.A.: Efficient multicriterial optimization based on intensive reuse of search information. J Glob Optim. \textbf{71}(1), 73--90 (2018). \doi{10.1007/s10898-018-0624-3}
\bibitem{c11}	Strongin, R., Sergeyev, Ya.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2nd ed. 2013, 3rd ed. 2014).
\bibitem{c12}	Sergeyev Y.D., Strongin R.G., Lera D.: Introduction to global optimization exploiting space-filling curves. Springer. (2013).
\bibitem{c13}	Zhigljavsky, A., {\v Z}ilinskas, A.: Stochastic Global Optimization. Springer, Berlin. (2008).
\bibitem{c14}	Locatelli, M., Schoen, F. Global Optimization: Theory, Algorithms, and Applications. SIAM. (2013).
\bibitem{c15}	Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. Princeton Univer-sity Press. (2016).

\end{thebibliography}
\end{document}

### 2019-11-21 2020/PAVT_Nested_Full/pavt_paper.tex
New experimental results

@@
 %
-\title{ Combining local and global methods in a parallel nested optimization scheme 
+\title{ Использование локального и глобального поиска в параллельной схеме вложенной оптимизации
+%Combining local and global search in a parallel nested optimization scheme 
 \thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
@@
 %
-\author{Konstantin Barkalov\letter\orcidID{0000-0001-5273-2471} \and Ilya Lebedev\orcidID{0000-0002-8736-0652} \and Maria Kocheganova \orcidID{0000-0002-4722-6299}}
+\author{Konstantin Barkalov\letter\orcidID{0000-0001-5273-2471} \and Ilya Lebedev\orcidID{0000-0002-8736-0652} \and Maria Kocheganova \orcidID{0000-0002-4722-6299} \and Victor Gergel \orcidID{0000-0002-4013-2329}}
 %
@@
 Предложена параллельная схема вычислений, которая учитывает данную особенность.
-Новая схема основана на идее рекурсивной оптимизации, когда на верхнем уровне рекурсии проводится оптимизации по параметрам, влияющим глобально (с помощью алгоритма глобального поиска), а на нижнем -- решение задач локальной оптимизации (с помощью локальных методов).
+Новая схема основана на идее рекурсивной оптимизации, когда на верхнем уровне рекурсии проводится оптимизации по параметрам, влияющим глобально (с помощью алгоритма глобального поиска), а на нижнем -- решение задач локальной оптимизации (с помощью локального метода).
 Возникающие при этом локальные подзадачи не будут оказывать влияния друг на друга, и их решение можно проводить параллельно.
@@
 \center
-\begin{minipage}{0.45\linewidth}
+\begin{minipage}{0.48\linewidth}
 \center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (a)}
 \end{minipage}
-\begin{minipage}{0.45\linewidth}
+\begin{minipage}{0.48\linewidth}
 \center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (b)}
@@
 Various modifications of this algorithm and the corresponding theory of
-convergence are given in \cite{Strongin2000,Sergeyev2013}.
+convergence are given in \cite{Strongin2000,Sergeyev2013,Barkalov2018,Strongin2018}.
 
@@
 В задачах локальной оптимизации, возникающих в приложениях, априорная информация о функциях бывает достаточно ограниченной.
-Например, может предполагаться лишь локальный характер зависимости функции от параметров, тогда как значения градиента (а, тем более, и гессиана) являются неизвестными.
+Например, может предполагаться лишь локальный характер зависимости функции от параметров, тогда как значения градиента (а, тем более, гессиана) являются неизвестными.
 В этом случае применяются методы прямого поиска, которые не используют каких-либо предположений о гладкости минимизируемой функции.
@@
 
-%Сравнение разных локальных методов
-%Эффективность распараллеливания
+Схема параллельных вычислений, описанная в предыдущих разделах, была реализована в программной системе Globalizer, разрабатываемой в ННГУ им. Н.И. Лобачевского \cite{globalizerSystem,Sysoyev2017}. 
+%Дать характеристики системы, указать поддержваемые технологии и компиляторы
+Система разработана на языке C++, ...
 
-Исследование масштабируемости предложенного параллельного алгоритма будем проводить, решая серию из 100 тестовых задач, в которых можно выделить параметры, влияющие локально.
-С целью имитации прикладных задач, обладающих указанным свойством, будем использовать функции вида
-\[
+%Дать технические характеристики суперкомпьютера
+Вычислительные эксперименты проводились на суперкомпьютере . 
+Узел суперкомпьютера располагает ...
+
+С целью имитации прикладных задач, в которых можно выделить параметры, влияющие локально, при проведении экспериментов использовались тестовые функции вида 
+\begin{equation}\label{test_problem}
 f(x,y) = G(y)+R(x),
-\]
+\end{equation}
 где $G(y)$ -- многоэкстремальная функция размерности $N=2$, а $R(x)$ -- унимодальная функция размертности $N \gg 2$.
-В качестве многоэкстрмальной части задачи будем рассматривать функции 
+В качестве многоэкстрмальной части задачи рассматривались функции 
 \begin{eqnarray} \nonumber \label{vagris}
@@
 Данный класс функций часто используется для тестирования алгоритмов глобальной оптимизации.  
-В качестве локальной части задачи будем использовать модификацию функции Розенброка 
+В качестве локальной части задачи использовалась модифицированная функция Розенброка 
 \[
@@
 \]
-размерности $N-2$, в которой точка экстремума смещается случайным образом в области поиска с помощью линейного преобразования координат.
+в которой точка экстремума смещалась случайным образом в области поиска с помощью линейного преобразования координат.
 
@@
 \center
-\begin{minipage}{0.45\linewidth}
+\begin{minipage}{0.48\linewidth}
 \center{\includegraphics[width=1.0\linewidth]{grishagin.png} \\ (a)}
 \end{minipage}
-\begin{minipage}{0.45\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{grishagin.png} \\ (b)}
+\begin{minipage}{0.48\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{rosenbrock.png} \\ (b)}
 \end{minipage}
@@
 
-\section{Conclusion}
+На Fig.~\ref{fig_level} отмечены также точки испытаний, выполненных global search algorithm and Hooke--Jeeves method при решении соответствующих задач.
+Расположение точек испытаний наглядно иллюстрирует как многоэкстремальность функции $G(y)$ (точки испытаний образуют неравномерную сетку в области поиска), так и овражность функции $R(x)$ (точки испытаний вытянуты вдоль дна оврага).
+
+В первой серии экспериментов были решены 100 задач вида (\ref{test_problem}), в которых размерность локальных подзадач равнялась $N=50$.
+Задача считалась решенной, если в процессе минимизации функции $G(y)$ алгоритм глобального поиска сгенерировал точку очередного испытания $y^k$ в $\epsilon$-окрестности глобального минимума $y^*$, т.е.
+$\left\|y^* - y^k\right\|\leq\epsilon$, где $\epsilon = 10^{-2}$.
+При этом точность поиска минимума при минимизации унимодальной функции $R(x)$ выбиралась в 1000 раз меньше, т.е. $\delta = 10^{-5}$.
+В алгоритме глобального поиска использовался параметр $r=3.0$ из (\ref{Rule_R}) и параметр построения развертки $m=10$; использованный локальный метод не имел дополнительных параметров.
+
+%Дать характеристики параллельного запуска - число узлов, процессов, процессоров и т.п.
+
+В table \ref{tab1} отражено среднее число итераций $K_{av}$, выполненных алгоритмом глобального поиска при минимизации функции $G(y)$, среднее время $T_{av}$ решения задачи в целом, а также ускорение по времени $S$  в зависимости от числа $p$ используемых MPI-процессов.
+
+\begin{table}
+\centering
+\caption{Результаты решения серии задач вида}\label{tab1}
+\begin{tabular}{cccc}
+\hline\noalign{\smallskip}
+ $\;\;\;p\;\;\;$  &  $\;\;\;K_{av}\;\;\;$ &  $\;\;\;T_{av}\;\;\;$ & $\;\;\;S\;\;\;$ \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+ 1  & 349.4  & 9.14 & ---  \\
+ 16 & 25.6   & 0.70 & 13.2 \\
+ 32 & 12.9   & 0.37 & 24.6 \\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
+%Привести описание экспериментов на задаче размерности 100+
+
+\section{Conclusion} %Заключение надо поправить!!!
 
 Результаты проведенных экспериментов на серии задач оптимизации, в которых можно выделить многоэкстремальную и унимодальную части задачи, показывают, что предложенная в работе схема организации параллельных вычислений обладает хорошим потенциалом параллелизма.
-В частности, при решении серии из 100 задач с числом переменных 52 (из которых две переменных оказывали глобальное влияние на целевую функцию, а пятьдесят влияли локально) получено линейное ускорение (в среднем).
-А решении задач с числом локально влияющих переменных $N=100$ было получено ускорение ...
-Проведенное исследование позволяет с уверенностью говорить о решении существенно многомерных задач оптимизации указанного класса.
+В частности, при решении серии из 100 задач с общим числом переменных 52 (из которых две переменных оказывали глобальное влияние на целевую функцию, а пятьдесят влияли локально) получено линейное ускорение (в среднем).
+
+А при решении задач с числом локально влияющих переменных $N=100$ было получено ускорение ...
+
+%Проведенное исследование позволяет с уверенностью говорить о решении существенно многомерных задач оптимизации указанного класса.
 

### 2019-11-21 2020/PAVT_Nested_Full/pavt_paper.tex
Some corrections in first section

@@
 %
-\title{ Использование локального и глобального поиска в параллельной схеме вложенной оптимизации
+\title{ Комбинированное использование локального и глобального поиска в параллельной схеме вложенной оптимизации 
 %Combining local and global search in a parallel nested optimization scheme 
@@
 В статье рассматриваются задачи глобальной оптимизации и численные методы их решения.
-О характере зависимости целевой функции от ее параметров делается предположение, что она является многоэкстремальной лишь по некоторым из переменных, а зависимость от остальных параметров носит локальный характер.
-Задачи такого типа могут возникать при идентификации математических моделей по результатам экспериментов.
+О характере зависимости целевой функции от ее параметров делается предположение, что она является многоэкстремальной лишь по некоторым из переменных, а зависимость от остальных носит локальный характер.
+Задачи такого типа могут возникать при идентификации неизвестных параметров математических моделей по результатам экспериментов.
 Предложена параллельная схема вычислений, которая учитывает данную особенность.
-Новая схема основана на идее рекурсивной оптимизации, когда на верхнем уровне рекурсии проводится оптимизации по параметрам, влияющим глобально (с помощью алгоритма глобального поиска), а на нижнем -- решение задач локальной оптимизации (с помощью локального метода).
+Новая схема основана на идее рекурсивной оптимизации, когда на верхнем уровне рекурсии проводится оптимизации по переменным, влияющим глобально (с помощью алгоритма глобального поиска), а на нижнем -- решение задач локальной оптимизации (с помощью локального метода).
 Возникающие при этом локальные подзадачи не будут оказывать влияния друг на друга, и их решение можно проводить параллельно.
@@
 
-Классическим стало применение методов глобальной оптимизации для идентификации параметров моделей по данным экспериментов.
+Классическим стало применение методов глобальной оптимизации для идентификации параметров математических моделей по данным экспериментов.
 В задачах такого вида требуется провести поиск значений неизвестных параметров модели, при которых результаты расчетов по модели  близки к результатам, полученным экспериментально.
@@
 В задачах подобной размерности регулярные методы поиска глобального решения не могут быть применены из-за чрезвычайно больших вычислительных затрат на покрытие области точками испытаний.
-Это остается справедливым даже в случае использования эффективных алгоритмов (например, \cite{Paulavicius2011,Evtushenko2009,Jones2009}), строящих существенно неравномерные покрытия области .
+Это остается справедливым даже в случае использования эффективных алгоритмов (например, \cite{Paulavicius2011,Evtushenko2009,Jones2009}), строящих существенно неравномерные покрытия.
 
-Однако в таких задачах часто можно указать параметры, которые будут оказывать наибольшее влияние на результат.
+Однако в задачах идентификации моделей часто можно указать параметры, которые будут оказывать наибольшее влияние на результат.
 Число таких параметров, как правило, является небольшим.
@@
 Описана общая схема организации параллельных вычислений.
-Section 3 посвящен описанию параллельного алгоритма глобального поиска с использованием с использованием space-filling curves. Приведена формулировка параллельного алгоритма, пример его работы.
+Section 3 посвящен описанию параллельного алгоритма глобального поиска с использованием с использованием space-filling curves.
+Приведена вычислительная схема параллельного алгоритма, описаны его основные свойства.
 Section 4 обсуждаются методы решения локальных подзадач в рамках схемы рекурсивной оптимизации. Дается кратное описание свойств задач поиска локального экстремума и известных методов их решения. 
-Section 5 содержит результаты численных экспериментов. Здесь производится оценка масшабируемости предложенного параллельного алгоритма при решении серии тестовых задач.  
+Section 5 содержит результаты численных экспериментов. Здесь производится оценка масшабируемости предложенной схемы параллельных вычислений при решении серии тестовых задач.  
 Section 6 concludes the paper.
@@
 \end{equation}
-В соответствии с данной схемой вычисление одного значения функции $\varphi (y)$ (данный процесс будем называть search trial) подразумевает решение унимодальной задачи (\ref{local_problem}), которое может быть выполнено одним из методов локальной оптимизации.
+В соответствии с (\ref{global_problem}) вычисление одного значения функции $\varphi (y)$ (данный процесс будем называть search trial) подразумевает решение унимодальной задачи (\ref{local_problem}), которое может быть выполнено одним из методов локальной оптимизации.
 Эффективность и проработанность методов локальной оптимизации позволяет решать подзадачи (\ref{local_problem}) с точностью, значительно превышающей точность методов глобальной оптимизации.
@@
 \center{\includegraphics[width=1.0\linewidth]{MPI_Tree.pdf}}
-\caption{Дерево подзадача в параллельной схеме вложенной оптимизации}
+\caption{Дерево подзадач в параллельной схеме вложенной оптимизации}
 \label{MPI_tree}
@@
 Каждая подзадача решается в отдельном процессе, обмен данными между процессами может быть организован с помощью MPI.
-Пересылки данных между процессами будут минимальные: требуется передать от корня к потомкам координаты точек $y^1, ..., y^p$, и получить обратно найденные минимальные значения
+Пересылки данных между процессами будут минимальные: требуется передать от корня к потомкам координаты точек $y^1, ..., y^p$ и получить обратно найденные минимальные значения
 \[
@@
 
-%связка
-Правила алгоритма глобального поиска, в котором на каждой итерации проводится параллельно $p$ испытаний, состоят в следующем.
+В соответствии с \cite{Strongin2000,Strongin2018} правила алгоритма глобального поиска, в котором на каждой итерации проводится параллельно $p$ испытаний, состоят в следующем.
 На первой итерации метода испытания проводятся параллельно в двух граничных точках $x^1 = 0, \; x^2 = 1$, а также в $(p-2)$ произвольных внутренних точках отрезка $[0,1]$, т.е.
@@
 
+%English begin
 Suppose $n\geq 1$  iterations of the method have been carried out, in which trials were performed at $k=k(n)$ points $x^i,1\leq i \leq k$.
@@
 convergence are given in \cite{Strongin2000,Sergeyev2013,Barkalov2018,Strongin2018}.
+%English end
 

### 2019-11-21 2020/PAVT_Nested_Full/pavt_paper.tex
Some minor corrections
Preparation for translation

@@
 Предложена параллельная схема вычислений, которая учитывает данную особенность.
-Новая схема основана на идее рекурсивной оптимизации, когда на верхнем уровне рекурсии проводится оптимизации по переменным, влияющим глобально (с помощью алгоритма глобального поиска), а на нижнем -- решение задач локальной оптимизации (с помощью локального метода).
+Новая схема основана на идее nested (recursive) optimization, когда на верхнем уровне рекурсии проводится оптимизации по переменным, влияющим глобально (с помощью метода глобальной оптимизации), а на нижнем -- решение задач локальной оптимизации (с помощью локального метода).
 Возникающие при этом локальные подзадачи не будут оказывать влияния друг на друга, и их решение можно проводить параллельно.
@@
 по ним функция является одноэкстремальной.
-В этом случае решение задачи можно организовать по схеме вложенной (рекурсивной) оптимизации.
+В этом случае решение задачи можно организовать по схеме nested (recursive) optimization.
 Решение  многоэкстремальной подзадачи, для которой требуется использовать сложные алгоритмы глобальной оптимизации, будет проводиться на верхнем уровне рекурсии.
@@
 
-Будем предполагать, что при любом фиксированном наборе значений $\overline{x}$ функция $f(\overline{x},y)$ является многоэкстремальной и удовлетворяет условию Липшица
+Будем предполагать, что при любом фиксированном наборе значений $\overline{x}$ функция $f(\overline{x},y)$ является многоэкстремальной и удовлетворяет Lipschitz condition
 \[
@@
 \varphi(y^*) = \min_{y\in D}  \varphi (y)
-%\min \left\{ f(x,y): x\in S, y \in D \right\} = \min_{y\in D}  \left\{ \min_{x\in S} f(x,y) \right\}.
 \end{equation}
@@
 Н.И.
-Лобачевского был предложен \cite{Barkalov2018,Strongin2018,globalizerSystem} эффективный параллельный алгоритм глобальной оптимизации, предназначенный для решения задач вида (\ref{global_problem}) .
+Лобачевского был разработан \cite{Barkalov2018,Strongin2018,globalizerSystem} эффективный параллельный алгоритм глобальной оптимизации, предназначенный для решения задач вида (\ref{global_problem}) .
 Основная идея распараллеливания состоит в такой организации вычислений, при которой параллельно проводятся несколько испытаний.
-Этот подход характеризуется эффективностью (распараллеливается именно та часть вычислительного процесса, в котором выполняется основной объем вычислений) и общностью (применим для широкого класса алгоритмов глобальной оптимизации).
+Этот подход характеризуется эффективностью (распараллеливается именно та часть вычислительного процесса, в котором выполняется основной объем вычислений) и общностью (применим для широкого класса алгоритмов).
 
-В разработанных методах используется оригинальный прием снижения размерности решаемой задачи. 
+В разработанном алгоритме глобального поиска используется оригинальный прием снижения размерности решаемой задачи. 
 Редукция размерности основана на фундаментальном факте, согласно которому $N$-мерный гиперпараллелепипед $D$ и отрезок $[0,1]$ вещественной оси являются равномощными множествами, и отрезок $[0,1]$ может быть непрерывно отображен в $D$ с помощью Peano curve $y(x)$, i.e.
@@
 \begin{equation}\label{1d_problem}
-\varphi(y (x^*)) = \min \left\{ \varphi (y(x)) : x\in [0,1] \right\}.
+\varphi(y (x^*)) = \min \left\{ \varphi (y(x)) : x\in [0,1] \right\},
 \end{equation}
+where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
+\[
+\left|\varphi(y(x'))-\varphi(y(x''))\right|\leq H\left|x'-x''\right|^{1/N}
+\]
+with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation
+$ H=2 L \sqrt{N+3}$ and $y(x)$ is the Peano curve from $[0,1]$ onto $D$.
 
-The algorithms for the numerical construction of the Peano curve approximations (called evolvents)
+The algorithms for the numerical construction of the Peano curve approximations (called \textit{evolvents})
 are given in \cite{Strongin2000,Sergeyev2013}.
@@
 
-%Параллельный алгоритм - пример работы? Нужен ли?
-
 \section{Local search algorithms}
-
-%Локальные задачи и локальные методы
 
@@
 Использование методов прямого поиска позволяет в этих условиях добиться лучших результатов.
-
-%Стандартные библиотеки - будем использовать одну из них
 

### 2019-11-21 2020/PAVT_Poster/poster-ru.tex
Add useful references

@@
 
-\bibitem{eremin}
-Ерёмин~И.И. Фейеровские методы для задач выпуклой и линейной оптимизации. 
-Челябинск: Изд-во ЮУрГУ, 2009. 200~с.
+\bibitem{Strongin91}
+Стронгин~Р.Г. Поиск глобального оптимума. М.: Знание, 1990. 48~с.
 
-\bibitem{sokolinsky}
-Соколинский~Л.Б. Организация параллельного выполнения запросов в многопроцессорной 
-машине баз данных с иерархической архитектурой // Программирование. 2001. №~6. С.~13--29.
+\bibitem{Grishagin07}
+Городецкий~С.Ю., Гришагин~В.А. Нелинейное программирование и многоэкстремальная оптимизация. Н.~Новгород: Изд-во ННГУ, 2007. 489~с.
 
-\bibitem{stonebraker}
-Stonebraker~M., Kemnitz~G. The POSTGRES Next-generation Database Management System // Communications of the ACM. 1991. Vol.~34, No.~10. P.~78--92.
+\bibitem{Strongin13}
+Стронгин~Р.Г., Гергель~В.П., Гришагин~В.A., Баркалов~К.А. Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013. 280~с.
+
+\bibitem{Barkalov14}
+Баркалов~К.А.
+Использование параллельных характеристических алгоритмов для решения многомерных задач глобальной оптимизации  // Вестник ЮУрГУ.
+Серия: Вычислительная математика и информатика.
+2014.
+Т.~3, №~4.
+С.~116--123.
+
+\bibitem{Barkalov16}
+Баркалов~К.А., Лебедев~И.Г., Соврасов~В.В., Сысоев~А.В.
+Реализация параллельного алгоритма поиска глобального экстремума функции на Intel Xeon Phi // Вычислительные методы и программирование.
+2016.
+Т.~17, №~1.
+С.~101--110.
+
+%\bibitem{sokolinsky}
+%Соколинский~Л.Б. Организация параллельного выполнения запросов в многопроцессорной машине баз данных с иерархической архитектурой // Программирование. 2001. №~6. С.~13--29.
+
+%\bibitem{stonebraker}
+%Stonebraker~M., Kemnitz~G.
+The POSTGRES Next-generation Database Management System // Communications of the ACM.
+1991.
+Vol.~34, No.~10.
+P.~78--92.
 DOI:~\href{http://dx.doi.org/10.1145/125223.125262}{10.1145/125223.125262}.

### 2019-11-23 2020/PAVT_Nested_Full/pavt_paper.tex
Добавил текст про харуктеристики сборки и кластера. Добавил абзац про запуск 100+

@@
 %Дать характеристики системы, указать поддержваемые технологии и компиляторы
-Система разработана на языке C++, ...
-
 %Дать технические характеристики суперкомпьютера
-Вычислительные эксперименты проводились на суперкомпьютере . 
-Узел суперкомпьютера располагает ...
+Для реализации параллельного алгоритма использовался GCC 5.5.0 и intel MPI 2017.
+Вычислительные эксперименты проводились на вычислительном кластере Ломоносов (НИВЦ МГУ имени М.В.
+Ломоносова).
+Узел кластера располагает процессором  Intel Xeon X5570 2.93GHz, 12 Gb RAM.
 
@@
 При этом точность поиска минимума при минимизации унимодальной функции $R(x)$ выбиралась в 1000 раз меньше, т.е. $\delta = 10^{-5}$.
-В алгоритме глобального поиска использовался параметр $r=3.0$ из (\ref{Rule_R}) и параметр построения развертки $m=10$; использованный локальный метод не имел дополнительных параметров.
+В алгоритме глобального поиска использовался параметр $r=4.5$ из (\ref{Rule_R}) и параметр построения развертки $m=10$; использованный локальный метод не имел дополнительных параметров.
 
+Задачи решались на 1, 4 и 8 узлах кластера, соответственно использовалось 2, 16 и 32 MPI процесса.
+Решение задачи верхнего уровня всегда производилось на одном процессе, при этом параллельно выполнялось от 1 до 31 параллельного вычисления значений нижнего уровня.
+Для решения задачи локальной оптимизации использовалось от 1 до 4 процессов на одном узле кластера.
 %Дать характеристики параллельного запуска - число узлов, процессов, процессоров и т.п.
@@
 \noalign{\smallskip}\hline\noalign{\smallskip}
- 1  & 349.4  & 9.14 & ---  \\
+ 2  & 349.4  & 9.14 & ---  \\
  16 & 25.6   & 0.70 & 13.2 \\
@@
 %Привести описание экспериментов на задаче размерности 100+
+Также был проведен эксперимент с решением задач вида (\ref{test_problem}), в которых размерность локальных подзадач равнялась $N=100$. Параметр запуска аналогичны предыдущей серии экспериментов.
+При использовании 1 MPI процесса для решения задач каждого уровня время решения составило 16.46 сек.
+, при использовании 15 процессов для решения локальных задач - 3.001, а при использовании 31 процесса для решения локальных задач - 0.86.
+Таким образом, ускорение составило 19,14 раза.
 

### 2019-11-24 2020/PAVT_Nested_Full/pavt_paper.tex
Before translation

--- 2019-11-21 2020/PAVT_Nested_Full/pavt_paper.tex
@@
 \title{ Комбинированное использование локального и глобального поиска в параллельной схеме вложенной оптимизации 
-%Combining local and global search in a parallel nested optimization scheme 
+%Combining local and global search in a parallel nested optimization scheme ? 
 \thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
@@
 
-В разработанном алгоритме глобального поиска используется оригинальный прием снижения размерности решаемой задачи. 
+В разработанном global search algorithm используется оригинальный прием снижения размерности решаемой задачи. 
 Редукция размерности основана на фундаментальном факте, согласно которому $N$-мерный гиперпараллелепипед $D$ и отрезок $[0,1]$ вещественной оси являются равномощными множествами, и отрезок $[0,1]$ может быть непрерывно отображен в $D$ с помощью Peano curve $y(x)$, i.e.
@@
 Схема параллельных вычислений, описанная в предыдущих разделах, была реализована в программной системе Globalizer, разрабатываемой в ННГУ им. Н.И. Лобачевского \cite{globalizerSystem,Sysoyev2017}.  
-%Дать характеристики системы, указать поддержваемые технологии и компиляторы
-Система разработана на языке C++, ...
-
-%Дать технические характеристики суперкомпьютера
-Вычислительные эксперименты проводились на суперкомпьютере . 
-Узел суперкомпьютера располагает ...
+Вычислительные эксперименты проводились на  Lomonosov supercomputer (Lomonosov Moscow State University); узел суперкомпьютера располагает two quad-core processors Intel Xeon X5570 2.93GHz and 12 Gb RAM.
+При сборке системы Globalizer для запуска на Lomonosov использовался  GCC 5.5.0 compiler and Intel MPI 2017.
 
@@
 
-В первой серии экспериментов были решены 100 задач вида (\ref{test_problem}), в которых размерность локальных подзадач равнялась $N=50$.
 Задача считалась решенной, если в процессе минимизации функции $G(y)$ алгоритм глобального поиска сгенерировал точку очередного испытания $y^k$ в $\epsilon$-окрестности глобального минимума $y^*$, т.е.
@@
 При этом точность поиска минимума при минимизации унимодальной функции $R(x)$ выбиралась в 1000 раз меньше, т.е. $\delta = 10^{-5}$.
-В алгоритме глобального поиска использовался параметр $r=3.0$ из (\ref{Rule_R}) и параметр построения развертки $m=10$; использованный локальный метод не имел дополнительных параметров.
+В алгоритме глобального поиска использовался параметр $r=4.5$ из (\ref{Rule_R}) и параметр построения развертки $m=10$; использованный локальный метод не имел дополнительных параметров.
 
-%Дать характеристики параллельного запуска - число узлов, процессов, процессоров и т.п.
+В первой серии экспериментов были решены 100 задач вида (\ref{test_problem}), в которых размерность локальных подзадач равнялась $N=50$.
+Задачи решались на 1, 4 и 8 узлах кластера на которых было задействовано от 2 до 32 MPI-процессов.
+В соответствии со схемой распараллеливания, представленной на Fig.~\ref{MPI_tree}, решение задачи глобальной оптимизации на верхнем уровне рекурсии проводится в одном (главном) процессе, который инициирует параллельное решение локальных задач задач нижнего уровня (от 1 до 31 задачи соответственно).
+Таким образом, минимально возможное число задействованных процессов равно 2: по одному на каждом уровне рекурсии.
 
@@
 \centering
-\caption{Результаты решения серии задач вида}\label{tab1}
+\caption{Результаты решения серии задач размерности $N=52$}\label{tab1}
 \begin{tabular}{cccc}
@@
 
-%Привести описание экспериментов на задаче размерности 100+
+Во второй серии экспериментов были решены 10 задач вида (\ref{test_problem}), в которых размерность локальных подзадач равнялась $N=100$.
+Параметр запуска аналогичны предыдущей серии экспериментов.
+Результаты решения приведены в table \ref{tab2}.
 
-\section{Conclusion} %Заключение надо поправить!!!
+\begin{table}
+\centering
+\caption{Результаты решения серии задач размерности $N=102$}\label{tab2}
+\begin{tabular}{cccc}
+\hline\noalign{\smallskip}
+ $\;\;\;p\;\;\;$  &  $\;\;\;K_{av}\;\;\;$ &  $\;\;\;T_{av}\;\;\;$ & $\;\;\;S\;\;\;$ \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+ 2  & 299.1  & 41.19 & ---  \\
+ 16 & 22.0   & 2.92 & 14.1 \\
+ 32 & 11.7   & 1.50 & 27.4 \\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
+\section{Conclusion}
 
 Результаты проведенных экспериментов на серии задач оптимизации, в которых можно выделить многоэкстремальную и унимодальную части задачи, показывают, что предложенная в работе схема организации параллельных вычислений обладает хорошим потенциалом параллелизма.
-В частности, при решении серии из 100 задач с общим числом переменных 52 (из которых две переменных оказывали глобальное влияние на целевую функцию, а пятьдесят влияли локально) получено линейное ускорение (в среднем).
+В частности, при решении серии из 100 задач с общим числом переменных 52 (из которых две переменных оказывали глобальное влияние на целевую функцию, а пятьдесят влияли локально) получено ускорение 24.6 при использовании 32 процессов.
+В случае решения задач с более сложными локальными подзадачами (с числом локальных переменных $N=100$) ускорение увеличивается вплоть до 27.4 при использовании 32 процессов.
 
-А при решении задач с числом локально влияющих переменных $N=100$ было получено ускорение ...
-
-%Проведенное исследование позволяет с уверенностью говорить о решении существенно многомерных задач оптимизации указанного класса.
-
-%Направлением дальнейших исследований может являться как расширение числа используемых локальных и глобальных методов, так и расширения круга рассматриваемых задач, включая задачи оптимизации с ограничениями, а также задачи многокритериальной оптимизации.
+Разработанная схема параллельных вычислений может применяться при идентификации сложных математических моделей, которые характеризуются большим числом (десятки и сотни) неизвестных параметров.
 

### 2019-11-28 2020/PAVT_Nested_Full/pavt_paper.tex
After translation

--- 2019-11-21 2020/PAVT_Nested_Full/pavt_paper.tex
@@
-%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%% author.tex 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%% Springer 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %
-% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% RECOMMENDED 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%
 %
@@
 
-\usepackage[utf8]{inputenc}
-\usepackage[english, russian]{babel} % Русские и английские переносы
+%\usepackage[utf8]{inputenc}
+%\usepackage[english, russian]{babel} % Русские и английские переносы
 
@@
 %
-\title{ Комбинированное использование локального и глобального поиска в параллельной схеме вложенной оптимизации 
-%Combining local and global search in a parallel nested optimization scheme 
+\title{Combining of local and global search in the parallel nested optimization scheme 
 \thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
@@
 %
-\titlerunning{ABBR}  % abbreviated title (for running head)
+\titlerunning{Combining of local and global search}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 %
-\author{Konstantin Barkalov\letter\orcidID{0000-0001-5273-2471} \and Ilya Lebedev\orcidID{0000-0002-8736-0652} \and Maria Kocheganova \orcidID{0000-0002-4722-6299} \and Victor Gergel \orcidID{0000-0002-4013-2329}}
+\author{Konstantin Barkalov\letter\orcidID{0000-0001-5273-2471} \and Ilya 
+Lebedev\orcidID{0000-0002-8736-0652} \and Maria Kocheganova \orcidID{0000-0002-4722-6299} 
+\and Victor Gergel \orcidID{0000-0002-4013-2329}}
 %
@@
 
-В статье рассматриваются задачи глобальной оптимизации и численные методы их решения.
-О характере зависимости целевой функции от ее параметров делается предположение, что она является многоэкстремальной лишь по некоторым из переменных, а зависимость от остальных носит локальный характер.
-Задачи такого типа могут возникать при идентификации неизвестных параметров математических моделей по результатам экспериментов.
-Предложена параллельная схема вычислений, которая учитывает данную особенность.
-Новая схема основана на идее nested (recursive) optimization, когда на верхнем уровне рекурсии проводится оптимизации по переменным, влияющим глобально (с помощью метода глобальной оптимизации), а на нижнем -- решение задач локальной оптимизации (с помощью локального метода).
-Возникающие при этом локальные подзадачи не будут оказывать влияния друг на друга, и их решение можно проводить параллельно.
-Проведены численные эксперименты на нескольких сотнях тестовых задач, подтверждающие эффективность предложенной схемы параллельных вычислений.
+In the present paper, the global optimization problems and the numerical methods of solving these 
+ones are considered. 
+Regarding the character of the dependence of the objective function on its parameters, an assumption 
+is made that it is multiextremal with respect to selected variables only, and the dependence on the rest 
+ones has a local character. 
+The problems of this type may arise in the identification of the unknown parameters of the 
+mathematical models on the base of the results of experiments. 
+A parallel computational scheme, which accounts for this feature has been proposed.
+The novel scheme is based on the idea of the nested (recursive) optimization when at the upper 
+recursion level the optimization with respect to the globally-meaning variables is performed (using 
+the global optimization method), and at the lower one the local optimization problems are solved (using 
+the local method). The generated local subproblems will not affect each other, and the solving of these 
+ones can be performed in parallel. The numerical experiments have been conducted on several 
+hundred test problems confirming the efficiency of the proposed scheme of parallel computations.
 
-\keywords{Global optimization $\cdot$ Local optimization $\cdot$ Recursive optimization scheme $\cdot$ Parallel algorithms}
+\keywords{Global optimization $\cdot$ Local optimization $\cdot$ Recursive optimization scheme 
+$\cdot$ Parallel algorithms}
 \end{abstract}
@@
 
-Задачи глобальной оптимизации часто возникают в различных областях современной науки и техники.
-Например,  актуальными являются проблемы поиска конфигураций различных химических соединений, соответствующих минимуму энергии взаимодействия \cite{Posypkin2014}.
-В частности, такие задачи возникают при разработке новых лекарственных препаратов \cite{Sulimov}.
-Задачи глобальной оптимизации возникают и во многих других приложениях (см.
-обзор  \cite{Pinter2006}).
+The global optimization problems arise in various fields of modern science and technology often. For 
+example, the problems of finding the configurations of various chemical compounds corresponding to 
+the minimum interaction energy are relevant \cite{Posypkin2014}. In particular,  such problems arise 
+in the development of new medical drugs \cite{Sulimov}. The global optimization problems arise in 
+many other applications as well (see the review \cite{Pinter2006}).
 
-Классическим стало применение методов глобальной оптимизации для идентификации параметров математических моделей по данным экспериментов.
-В задачах такого вида требуется провести поиск значений неизвестных параметров модели, при которых результаты расчетов по модели  близки к результатам, полученным экспериментально.
+The application of the global optimization methods for the identification of the parameters of the 
+mathematical models based on the data of experiments became a classical one. In the problems of this 
+kind, it is required to carry out a search for the values of the unknown parameters of the model, at 
+which the results of computations according to the model are close to the results obtained 
+experimentally.
 
-Число параметров, которые требуется идентифицировать подобным образом, для сложных математических моделей может составлять десятки и сотни \cite{Nurislamova2016,Akhmadullina2017}.
-В задачах подобной размерности регулярные методы поиска глобального решения не могут быть применены из-за чрезвычайно больших вычислительных затрат на покрытие области точками испытаний.
-Это остается справедливым даже в случае использования эффективных алгоритмов (например, \cite{Paulavicius2011,Evtushenko2009,Jones2009}), строящих существенно неравномерные покрытия.
+The number of parameters, which are required to be identified in such a way may be tens and hundreds 
+in complex mathematical models \cite{Akhmadullina2017,Nurislamova2016}. In the problems of 
+such dimensionality, the regular methods of search for the global solutions cannot be applied because 
+of extremely large computational costs for covering the search domain by the trial points. This 
+remains to be true even in the case of the use of efficient algorithms (for example, 
+\cite{Evtushenko2009,Jones2009,Paulavicius2011}) constructing essentially nonuniform coverages.
 
-Однако в задачах идентификации моделей часто можно указать параметры, которые будут оказывать наибольшее влияние на результат.
-Число таких параметров, как правило, является небольшим.
-Остальные параметры либо оказывают незначительное влияние (в этом случае многоэкстремальность можно не учитывать), либо зависимость от второй группы параметров носит локальный характер.
+However, in the problems of the model identification, the parameters, which will affect the result 
+mostly can be selected often. The number of such parameters is small as a rule. The rest parameters 
+either cause an insufficient effect (in this case, the multiextremality can be neglected) or the 
+dependence on the second group of parameters have a local character.
 
-Таким образом, актуальной является разработка параллельных алгоритмов минимизации существенно многомерных функций (сотни переменных), в которых учитывается разный характер зависимости целевой функции от разных групп параметров.
-А именно, предполагается, что целевая функция является многоэкстремальной лишь по части переменных.
-Остальные переменные влияют локально, т.е.
-по ним функция является одноэкстремальной.
-В этом случае решение задачи можно организовать по схеме nested (recursive) optimization.
-Решение  многоэкстремальной подзадачи, для которой требуется использовать сложные алгоритмы глобальной оптимизации, будет проводиться на верхнем уровне рекурсии.
-Одноэкстремальные подзадачи (каждая из которых соответствует фиксированному набору значений первой части параметров) будут решаться на нижнем уровне.
-Для решения одноэкстремальных задач можно применять эффективные методы локальной оптимизации, линейного программирования или же линейной алгебры (в зависимости от характера локальных подзадач).
+So far, the development of the parallel algorithms for the minimization of the essentially 
+multidimensional functions (hundreds of variables), which account for different character of 
+dependence of the objective function on different groups of parameters is relevant. Namely, the 
+objective function is assumed to be multiextremal with respect to a part of variables only. The rest 
+variables have a local effect i. e. the objective function is monoextremal with respect to these ones. 
+In this case, the solving of the problem can be organized according to the nested (recursive) 
+optimization scheme. The solving of the multiextremal subproblem, for which the use of complex 
+global search algorithms is required will be performed at the upper level of recursion.
+The monoextremal subproblems (each corresponding to a fixed set of values of the first part of 
+parameters) will be solved at the lower level. For solving the monoextremal problems, the efficient 
+methods of local optimization, linear programming, or linear algebra can be applied (subject to the 
+character of the local subproblems).
 
-Текст статьи, отражающей результаты проведенного исследования, построен следующим образом. 
-В Section 2 дано описание схемы вложенной оптимизации с учетом разных свойств функции на разных уровнях рекурсии.
-Приведен конкретный пример задачи указанного класса.
-Описана общая схема организации параллельных вычислений.
-Section 3 посвящен описанию параллельного алгоритма глобального поиска с использованием с использованием space-filling curves.
-Приведена вычислительная схема параллельного алгоритма, описаны его основные свойства.
-Section 4 обсуждаются методы решения локальных подзадач в рамках схемы рекурсивной оптимизации. Дается кратное описание свойств задач поиска локального экстремума и известных методов их решения. 
-Section 5 содержит результаты численных экспериментов. Здесь производится оценка масшабируемости предложенной схемы параллельных вычислений при решении серии тестовых задач.  
+The paper text describing the results of the present study, is organized as follows. 
+In Section 2, the description of the nested optimization scheme taking into account different properties 
+of the objective function at different nesting level is given. Particular example of a problem of this 
+class is presented. A general scheme of organization of the parallel computations is described. 
+Section 3 is devoted to the description of the parallel global search algorithm using the space-filling 
+curves. The computational scheme of the parallel algorithm is presented, its main properties are 
+described.
+In Section 4, the methods of solving the local subproblems within the framework of the nested 
+optimization scheme are discussed. A brief description of the properties of the local extremum search 
+problems and known methods of solving these ones is  given. 
+Section 5 contains the results of numerical experiment. Here, the evaluation of the scalability of the 
+proposed parallel computation scheme in solving a series of test problems is performed.  
 Section 6 concludes the paper.
@@
 
-Рассмотрим задачу оптимизации вида
+Let us consider an optimization problem of the kind
 \begin{eqnarray}\label{main_problem}
@@
 
-Будем предполагать, что при любом фиксированном наборе значений $\overline{x}$ функция $f(\overline{x},y)$ является многоэкстремальной и удовлетворяет Lipschitz condition
+Let us assume that at any fixed set of the values $\overline{x}$ the function $f(\overline{x},y)$ is a 
+multiextremal one and satisfies Lipschitz condition
 \[
@@
 with the constant $L$ unknown a priori.
-Одновременно с этим при любом фиксированном наборе значений $\overline{y}$  функция $f(x,\overline{y})$ является унимодальной, т.е.
-функция $f(x,\overline{y})$ имеет единственную точку минимума $x^*$ (зависящую, вообще говоря, от $\overline{y}$).
+At the same time, at any fixed set of values $\overline{y}$ the function $f(x,\overline{y})$ is a 
+unimodal one i.e. the function $f(x,\overline{y})$ has a single minimum point $x^*$ (depending, in 
+general, on $\overline{y}$).
 
-Учет подобной особенности рассматриваемого класса задач может существенно снизить вычислительную сложность поиска оптимума.
-В самом деле, в соответствии с известной схемой рекурсивной оптимизации \cite{Carr} решение исходной задачи можно свести к задаче поиска глобального минимума функции $\varphi(y)$
+Accounting for such a peculiarity of the considered problem class can reduce the computational costs 
+of the optimum search essentially. Indeed, according to known nested optimization scheme 
+\cite{Carr}, the solving of the initial problem can be reduced to the problem of finding the global 
+minimum of the function $\varphi(y)$
 \begin{equation}\label{global_problem}
@@
 \end{equation}
-где 
+where 
 \begin{equation}\label{local_problem}
@@
 \end{equation}
-В соответствии с (\ref{global_problem}) вычисление одного значения функции $\varphi (y)$ (данный процесс будем называть search trial) подразумевает решение унимодальной задачи (\ref{local_problem}), которое может быть выполнено одним из методов локальной оптимизации.
-Эффективность и проработанность методов локальной оптимизации позволяет решать подзадачи (\ref{local_problem}) с точностью, значительно превышающей точность методов глобальной оптимизации.
-Соответственно, можно рассматривать задачу (\ref{global_problem}) как задачу глобальной оптимизации, в которой значение функции может быть вычислено с высокой точностью, но данная операция является трудоемкой.
+According to (\ref{global_problem}), the computing of a single value of the function $\varphi (y)$ 
+(let us call this process a search trial) implies the solving of a unimodal problem (\ref{local_problem}), 
+which can be performed by one of the local search methods. The efficiency and sophistication of the 
+local optimization methods allow solving the subproblems (\ref{local_problem}) with the precision 
+exceeding the one of the global optimization methods essentially. Correspondingly, the problem 
+(\ref{global_problem}) can be considered as a global optimization problem, in which the objective 
+function value can be computed with high precision but this operation is computation-costly.
 
-Конкретным примером задачи, в которой наблюдается разный характер зависимостей от разных групп параметров, может являться следующая задача аппроксимации.
-
-Пусть в ходе эксперимента получены $m$ значений $u_j = u(t_j), 1 \leq j \leq m, $ функции $u(t)$, причем известно, что аналитическое выражение для  $u(t)$ имеет вид
+The following approximation problem can be considered as a particular example of a problem, in 
+which different character of dependence on different groups of parameters Takes place. 
+Let us assume $m$ values $u_j = u(t_j), 1 \leq j \leq m, $ of the function $u(t)$ to have been obtained 
+in the experiment and the analytical expression for $u(t)$ is known to take the form 
 \begin{equation}\label{ex_func}
@@
 \end{equation}
-В данном выражении $c=\left(c_1,\ldots,c_{2n+2}\right)$, $\omega=\left(\omega_0,\omega_1,\ldots,\omega_n\right)$ являются неизвестными параметрами, определяющими конкретную функцию $u(t)$.
+In this expression, $c=\left(c_1,\ldots,c_{2n+2}\right)$, 
+$\omega=\left(\omega_0,\omega_1,\ldots,\omega_n\right)$ are the unknown parameters defining 
+particular function $u(t)$.
 
-Введем меру отклонения функции $u(t,\omega,c)$ от экспериментальных данных как сумму квадратов
+Let us introduce a measure of response of the function $u(t,\omega,c)$ to the experiential data as a 
+sum of squares 
 \[
@@
 \]
-Now, following the idea of least squares fitting, it is possible to present the approximation problem as the problem of minimizing the objective function
+Now, following the idea of least squares fitting, it is possible to present the approximation problem as 
+the problem of minimizing the objective function
 \begin{equation}\label{ex_prob}
@@
 \end{equation}
-Решение $(\omega^*,c^*)$ данной задачи будет определять наилучшее приближение.
+The solution $(\omega^*,c^*)$ of this problem will define the best approximation.
 
-Задача (\ref{ex_prob}) может быть записана в рекурсивной форме
+The problem (\ref{ex_prob}) can be written in a recursive form 
 \begin{equation}\label{outer_prob}
@@
 \end{equation}
-Вложенная подзадача (\ref{inner_prob}) является классической линейной задачей наименьших квадратов, и ее решение can be obtained by solving a system of linear algebraic equations regarding the unknown $c$, which can be done, e.g., by Gaussian elimination.
-Одновременно с этим внешняя задача (\ref{outer_prob}) будет являться, вообще говоря, многоэкстремальной и ее решение потребует значительных вычислительных ресурсов.
+The nested subproblem (\ref{inner_prob}) is a classical linear least squares problem, and its solution 
+can be obtained by solving a system of linear algebraic equations regarding the unknown $c$, which 
+can be done e.g. by Gaussian elimination. At the same time, the external problem (\ref{outer_prob}) 
+will be, generally speaking, a multiextremal one, and its solving will require considerable 
+computational resources.
 
-Таким образом, для решения задачи (\ref{main_problem}) в рекурсивной схеме (\ref{global_problem}), (\ref{local_problem}) можно применять параллельный алгоритм глобального поиска для решения внешних подзадач (\ref{global_problem}).
-При этом на каждой итерации алгоритма будет порождаться множество независимых подзадач (\ref{local_problem}), решение которых может проводиться параллельно с помощью локальных методов.
-Общая схема организации вычислений с использованием нескольких процессов приведена на Fig.~\ref{MPI_tree}.
+So far, to solve the problem (\ref{main_problem}) in the nested scheme (\ref{global_problem}), 
+(\ref{local_problem}), one can apply parallel global search algorithm for solving the external 
+subproblems (\ref{global_problem}). At that, a set of independent subproblems (\ref{local_problem}) 
+will be generated at every iteration of the algorithm, the solving of which can be performed in parallel 
+using the local methods.
+A general scheme of the organization of the computations using several processors is presented in 
+Fig.~\ref{MPI_tree}.
 
@@
 \center{\includegraphics[width=1.0\linewidth]{MPI_Tree.pdf}}
-\caption{Дерево подзадач в параллельной схеме вложенной оптимизации}
+\caption{A tree of subproblems in the parallel nested optimization scheme}
 \label{MPI_tree}
@@
 
-Процессы параллельной программы будут образовывать дерево. 
-Корневой процесс будет решать задачу (\ref{global_problem}) и распределять подзадачи (\ref{local_problem}) между процессами-потомками.
-Каждая подзадача решается в отдельном процессе, обмен данными между процессами может быть организован с помощью MPI.
-Пересылки данных между процессами будут минимальные: требуется передать от корня к потомкам координаты точек $y^1, ..., y^p$ и получить обратно найденные минимальные значения
+The processes of the parallel program will form a tree. 
+The root process will solve the problem (\ref{global_problem}) and distribute the subproblems 
+(\ref{local_problem}) among the child processes. Each subproblem is solved by a separate process, 
+the data exchange between the processes can be organized using MPI. The data transmission between 
+the processes will be minimal: it is required to send the coordinates of points $y^1, ..., y^p$ from the 
+root to the child processes and to receive back the found minimal values 
 \[
@@
 
-Подробное обсуждение алгоритмов, используемых на разных уровнях дерева процессов, приведено в следующих разделах.
+A detailed discussion of the algorithms used at different levels of the process tree is given in the next 
+sections.
 
@@
 
-В ННГУ им.
-Н.И.
-Лобачевского был разработан \cite{Barkalov2018,Strongin2018,globalizerSystem} эффективный параллельный алгоритм глобальной оптимизации, предназначенный для решения задач вида (\ref{global_problem}) .
-Основная идея распараллеливания состоит в такой организации вычислений, при которой параллельно проводятся несколько испытаний.
-Этот подход характеризуется эффективностью (распараллеливается именно та часть вычислительного процесса, в котором выполняется основной объем вычислений) и общностью (применим для широкого класса алгоритмов).
+An efficient parallel global optimization algorithm intended for solving the problems of the kind 
+(\ref{global_problem}) has been developed in Lobachevskii University of Nizhnii Novgorod (UNN) 
+\cite{Barkalov2018,globalizerSystem,Strongin2018}. 
+The main idea of the parallelization consists in such of organization of computations that several trials 
+are performed in parallel. This approach is featured by high efficiency (the part of the computational 
+process is parallelized, in which the major amount of computations is performed) and generality (it is 
+applicable for a wide class of algorithms).
 
-В разработанном алгоритме глобального поиска используется оригинальный прием снижения размерности решаемой задачи. 
-Редукция размерности основана на фундаментальном факте, согласно которому $N$-мерный гиперпараллелепипед $D$ и отрезок $[0,1]$ вещественной оси являются равномощными множествами, и отрезок $[0,1]$ может быть непрерывно отображен в $D$ с помощью Peano curve $y(x)$, i.e.
-$D = \left\{y(x):0\leq x\leq 1\right\}$.
+In the developed global search algorithm, a novel method for the reduction of the dimensionality of 
+the problems to be solved is applied. 
+The reduction of the dimensionality is based on a fundamental fact, according to which an 
+$N$-dimensional hyperparallelepiped $D$ and the interval $[0,1]$ of the real axis are the 
+equal-power sets, and the interval $[0,1]$ can be mapped continuously into $D$ using the Peano 
+curve $y(x)$ i. e. $D = \left\{y(x):0\leq x\leq 1\right\}$.
 
-Используя этот факт, можно свести решение задачи минимизации многомерной функции $\varphi(y)$ к минимизации одномерной функции $\varphi(y(x))$
+Utilizing this fact, one can reduce the solving of the problem of minimization of a multidimensional 
+function $\varphi(y)$ to the minimization of a one-dimensional function $\varphi(y(x))$
 \begin{equation}\label{1d_problem}
@@
 
-The algorithms for the numerical construction of the Peano curve approximations (called \textit{evolvents})
-are given in \cite{Strongin2000,Sergeyev2013}.
-As an illustration, two evolvents are shown in Fig.~\ref{fig_evolvent}.
-Figure demonstrates that the precision of the evolvent is depend on the density level $m$ used in the construction.
+The algorithms for the numerical construction of the Peano curve approximations (called 
+\textit{evolvents})
+are given in \cite{Sergeyev2013,Strongin2000}. As an illustration, two evolvents are shown in 
+Fig.~\ref{fig_evolvent}. Figure demonstrates that the precision of the evolvent is depend on the 
+density level $m$ used in the construction.
 
@@
 
-Отметим, что многие известные алгоритмы глобальной оптимизации неявно основаны на идее редукции размерности и адаптации одномерных алгоритмов для решения многомерных задач \cite{Sergeyev2006,Zilinskas2008,Evtushenko2013}.
+It is worth noting that many well known global optimization algorithms are based implicitly on the 
+idea of the dimensionality reduction and adaptation of the one-dimensional algorithms for solving the 
+multidimensional problems \cite{Evtushenko2013,Sergeyev2006,Zilinskas2008}.
 
-В соответствии с \cite{Strongin2000,Strongin2018} правила алгоритма глобального поиска, в котором на каждой итерации проводится параллельно $p$ испытаний, состоят в следующем.
-На первой итерации метода испытания проводятся параллельно в двух граничных точках $x^1 = 0, \; x^2 = 1$, а также в $(p-2)$ произвольных внутренних точках отрезка $[0,1]$, т.е.
+According to \cite{Strongin2018,Strongin2000}, the rules of the global search algorithm, in which 
+$p$ trials are performed in parallel at every iteration, are as follows.
+At the first iteration of the method, the trials are performed in parallel in two boundary points $x^1 = 
+0, \; x^2 = 1$ as well as in $(p-2)$ arbitrary internal points of the interval $[0,1]$ i. e. 
 $x^i\in(0,1),3\leq i \leq p$.
@@
 %English begin
-Suppose $n\geq 1$  iterations of the method have been carried out, in which trials were performed at $k=k(n)$ points $x^i,1\leq i \leq k$.
-Then points $x^{k+1},\dots,x^{k+p}$  of search trials of the next iteration are determined according to the following rules.
+Suppose $n\geq 1$  iterations of the method have been carried out, in which trials were performed at 
+$k=k(n)$ points $x^i,1\leq i \leq k$. Then points $x^{k+1},\dots,x^{k+p}$  of search trials of the next 
+iteration are determined according to the following rules.
 
@@
 \end{equation}
-where $\Delta_i=(x_i-x_{i-1})^{1/N}$. The value $r > 1$ is parameter of the algorithm. An appropriate selection
+where $\Delta_i=(x_i-x_{i-1})^{1/N}$. The value $r > 1$ is parameter of the algorithm. An 
+appropriate selection
 of $r$ allows to consider the product $r \mu$ as an estimate
@@
 
-Rule 5. Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), \; 1\leq j\leq p$, calculated using the formulae
+Rule 5. Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), \; 1\leq j\leq p$, calculated 
+using the formulae
 \begin{equation}\label{Rule_X}
-x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - \frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N.
+x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - 
+\frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^
+N.
 \end{equation}
 
-The algorithm terminates if the condition $\Delta_{t_j}<\epsilon$ is satisfied at least for one number $t_j, 1 \leq j \leq p$ ; $\epsilon>0$ is the predefined accuracy.
+The algorithm terminates if the condition $\Delta_{t_j}<\epsilon$ is satisfied at least for one number 
+$t_j, 1 \leq j \leq p$ ; $\epsilon>0$ is the predefined accuracy.
 
-This method of parallelization has the following justification.
-The characteristics of intervals (\ref{Rule_R}) used in the method can be considered as probability measures of the global minimum point location in these intervals.
-Inequalities (\ref{Rule_Max}) arrange intervals according to their characteristics, and trials are carried out in parallel in the first $p$ intervals with the largest probabilities.
+This method of parallelization has the following justification. The characteristics of intervals 
+(\ref{Rule_R}) used in the method can be considered as probability measures of the global minimum 
+point location in these intervals. Inequalities (\ref{Rule_Max}) arrange intervals according to their 
+characteristics, and trials are carried out in parallel in the first $p$ intervals with the largest 
+probabilities.
 
 Various modifications of this algorithm and the corresponding theory of
-convergence are given in \cite{Strongin2000,Sergeyev2013,Barkalov2018,Strongin2018}.
+convergence are given in \cite{Barkalov2018,Sergeyev2013,Strongin2018,Strongin2000}.
 %English end
@@
 
-В данном разделе рассмотрим вопросы, связанные с решением подзадачи (\ref{local_problem}), которой (при обозначении $f(x) = f(x,\overline{y})$ для фиксированного значения $\overline{y} \in D$) соответствует задача на локальный экстремум
+In the present section, let us consider the issues related to the solving the subproblem 
+(\ref{local_problem}), which (in the notation $f(x) = f(x,\overline{y})$ for the fixed value 
+$\overline{y} \in D$) corresponds the problem of finding a local extremum
 
@@
 
-К настоящему времени разработано огромное количество разнообразных методов локальной оптимизации для задач вида (\ref{lp}).
-К их числу относятся, например, градиентные методы, ньютоновские и квазиньютоновские методы, методы сопряженных направлений.
-Большинство из них используют принцип локального спуска, когда метод последовательно на каждом шаге переходит к точкам с меньшими значениями целевой функции.
-Почти все эти методы могут быть представлены в виде итерационного соотношения
+To date, a huge number of various local optimization methods for the problems of the kind (\ref{lp}) 
+have been developed. Among these ones, there are, for example, the gradient methods, Newton and 
+quasi-Newton methods, the methods of conjugate directions, etc. The majority of these ones utilize 
+the principle of local descent, when the method passes to the points with smaller values of the 
+objective function at every step progressively. Almost all these methods cam be represented in the 
+form of the iterative relation 
 \[
@@
 \]
-где $x^k$ -- точки основных испытаний, состоящих в вычислении набора тех или иных локальных характеристик целевой функции $I^k=I(x^k)$ в точке $x^k$, $d^k$ -- направления смещения из точек $x^k$, вычисляемые по результатам основных испытаний, а $s^k$ -- коэффициенты, определяющие величины смещений вдоль выбранных направлений.
+where $x^k$ are the points of main trials consisting in computing a set of some local characteristics of 
+the objective function $I^k=I(x^k)$ at the point $x^k$, where $d^k$ are the directions of shift from 
+the points $x^k$ computed on the base of the results of main trials and $s^k$ are the coefficients 
+defining the magnitudes of the shifts along to the directions chosen.
 
-Для определения величин смещений $s^k$ вдоль направлений $d^k$ методы могут выполнять вспомогательные (рабочие) шаги.
-Это приводит к дополнительным измерениям локальных характеристик целевой функции вдоль направления $d^k$.
-Переходы от точек $x^k$ к точкам $x^{k+1}$ выполняются таким образом, чтобы обеспечить существенное убывание значений функции $f^k = f( x^k )$ в результате шага.
+In order to define the shift magnitudes $s^k$ along the directions $d^k$, the methods can perform 
+auxiliary (working) steps. This results in additional measuring of the local characteristics of the 
+objective function along the direction $d^k$. The transitions from the points $x^k$ to the points 
+$x^{k+1}$ is performed in such a way that an essential decrease of the value of the function $f^k = 
+f( x^k )$ is provided as a result of the step.
 
-В набор вычисляемых локальных характеристик $I^k=I(x^k)$ могут входить: значение функции $f^k = f( x^k )$, вектор градиента $\nabla f^k = \nabla f(x^k)$, матрица вторых производных (гессиан) $\Gamma_k=\Gamma^f(x^k)$.
-Какой именно набор характеристик измеряется -- зависит как от свойств решаемой задачи, так и от выбранного метода оптимизации.
+The set of the computed local characteristics $I^k=I(x^k)$ may include: the function values $f^k = 
+f( x^k )$, the vector of gradient $\nabla f^k = \nabla f(x^k)$, the matrix of the second derivatives 
+(Hessian) $\Gamma_k=\Gamma^f(x^k)$. Particular set of the characteristics to be measured depends 
+on the properties of the problem being solved as well as on the optimization method chosen.
 
-В задачах локальной оптимизации, возникающих в приложениях, априорная информация о функциях бывает достаточно ограниченной.
-Например, может предполагаться лишь локальный характер зависимости функции от параметров, тогда как значения градиента (а, тем более, гессиана) являются неизвестными.
-В этом случае применяются методы прямого поиска, которые не используют каких-либо предположений о гладкости минимизируемой функции.
-При поиске минимума эти методы измеряют только значения функции.
-Правила размещения точек итераций в них основываются на некоторых эвристических логических схемах.
+In the local optimization problems arising in the applications, the a priori information on the objective 
+functions is quite limited usually. For example, a local character of the dependence of the function on 
+the parameters may be suggested only whereas the gradient values (and, moreover, Hessian) are
+unknown. In this case, the direct search methods are applied, which don't utilize any suggestions on
+the smoothness of the minimized function. When searching for the minimum, these methods measure 
+the function values only. The rules of the placement of the iteration points in these ones is based on 
+some heuristic logical schemes.
 
-Одними из популярных методов прямого поиска являются Hooke--Jeeves method \cite{HookJeeves} and Nelder--Mead method \cite{NelderMead}.
-Несмотря на кажущуюся простоту и теоретическую необоснованность методов прямого поиска, они хорошо зарекомендовали себя в реальных расчетах.
-Это можно объяснить следующим образом.
-Многие методы гладкой оптимизации чрезвычайно чувствительны к наличию вычислительных ошибок в значениях функций, превращающих теоретически гладкую функцию в фактически негладкую.
-За счет этого в реальных расчетах они зачастую утрачивают те положительные свойства, которые для них обещает теория.
-Использование методов прямого поиска позволяет в этих условиях добиться лучших результатов.
+Hooke--Jeeves method \cite{HookJeeves} and Nelder--Mead method \cite{NelderMead} belong to 
+the popular methods of the direct search. In spite of apparent simplicity and theoretical 
+non-substantiation of the direct search methods, these ones are well established in the practical 
+computations. This can be explained as follows. Many methods of smooth optimization are very 
+sensitive to the computational errors in the function values transforming theoretically smooth function 
+into the non-smooth one actually. Because of this, in the practical computations these ones loose the 
+positive properties, which are promised by theory often. The use of the direct search methods allows 
+achieving better results in these conditions.
 
@@
 
-Схема параллельных вычислений, описанная в предыдущих разделах, была реализована в программной системе Globalizer, разрабатываемой в ННГУ им. Н.И. Лобачевского \cite{globalizerSystem,Sysoyev2017}. 
-%Дать характеристики системы, указать поддержваемые технологии и компиляторы
-Система разработана на языке C++, ...
+The scheme of parallel computations described in previous sections was implemented in the 
+Globalizer software system developed in UNN \cite{globalizerSystem,Sysoyev2017}.  
+The numerical experiments were carried out using Lomonosov supercomputer (Lomonosov Moscow 
+State University). Each supercomputer node included two quad-core processors Intel Xeon X5570 
+2.93GHz and 12 Gb RAM. When building the Globalizer system for running on the Lomonosov 
+supercomputer GCC 5.5.0 compiler and Intel MPI 2017 were used.
 
-%Дать технические характеристики суперкомпьютера
-Вычислительные эксперименты проводились на суперкомпьютере . 
-Узел суперкомпьютера располагает ...
-
-С целью имитации прикладных задач, в которых можно выделить параметры, влияющие локально, при проведении экспериментов использовались тестовые функции вида 
+In order to simulate the applied problems, in which the locally-affecting parameters can be selected, 
+the test function of the kind 
 \begin{equation}\label{test_problem}
@@
 \end{equation}
-где $G(y)$ -- многоэкстремальная функция размерности $N=2$, а $R(x)$ -- унимодальная функция размертности $N \gg 2$.
-В качестве многоэкстрмальной части задачи рассматривались функции 
+were used when conducting the experiments. Here $G(y)$ is a multiextremal function of the 
+dimensionality $N=2$ and $R(x)$ is a unimodal function of the dimensionality $N \gg 2$.
+As the multiextremal part of the problem, the functions 
 \begin{eqnarray} \nonumber \label{vagris}
-G(y)= -&\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2+\right. \\
-&\left.\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2},\\ \nonumber
+G(y)= 
+-&\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2+\right. \\
+&\left.\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2
+},\\ \nonumber
 \end{eqnarray}
-where
+were considered where
 \begin{eqnarray} \nonumber
@@
 \end{eqnarray}
-and coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$  are taken uniformly in the interval $[-1,1]$.
-Данный класс функций часто используется для тестирования алгоритмов глобальной оптимизации.  
-В качестве локальной части задачи использовалась модифицированная функция Розенброка 
+and the coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$ are the random numbers distributed uniformly 
+in the interval $[-1,1]$.
+The class of functions is used for testing the global optimization algorithms often. 
+As the local part of the problem, the modified Rosenbrock function 
 \[
-R(x)= \sum_{i=1}^{N-1}{\left[(1-x_i)^2+100(x_{i+1}-x_i^2)^2\right]}, -2 \leq x_i \leq 2 , 1\leq i\leq N,
+R(x)= \sum_{i=1}^{N-1}{\left[(1-x_i)^2+100(x_{i+1}-x_i^2)^2\right]}, -2 \leq x_i \leq 2 , 1\leq 
+i\leq N,
 \]
-в которой точка экстремума смещалась случайным образом в области поиска с помощью линейного преобразования координат.
+was used, in which the extremum point was shifted in the search domain randomly using the linear 
+coordinate transform.
 
-В качестве примера на Fig.~\ref{fig_level}(a),(b) приведены линии уровня одной пары функций $G(y)$ и $R(x)$.
-Данные функции являются сложными для соответствующих методов глобальной и локальной оптимизации, т.к.
-функции $G(y)$ являются существенно многоэкстремальными, а функции $R(x)$ обладают ярко выраженной овражной структурой.
+As an example, a contour plot of a pair of functions  $G(y)$ and $R(x)$ is presented in 
+Fig.~\ref{fig_level}(a),(b). These functions are the complex ones for corresponding global and local 
+optimization methods since the functions $G(y)$ are essentially multiextremal ones, and the functions 
+$R(x)$ have a ravine structure manifested clearly.
 
@@
 \end{minipage}
-\caption{Линии уровня функций (a) $G(y)$ и (b) $R(x)$}
+\caption{Contour plot of the functions (a) $G(y)$ and (b) $R(x)$}
 \label{fig_level}
@@
 
-На Fig.~\ref{fig_level} отмечены также точки испытаний, выполненных global search algorithm and Hooke--Jeeves method при решении соответствующих задач.
-Расположение точек испытаний наглядно иллюстрирует как многоэкстремальность функции $G(y)$ (точки испытаний образуют неравномерную сетку в области поиска), так и овражность функции $R(x)$ (точки испытаний вытянуты вдоль дна оврага).
+In Fig.~\ref{fig_level}, the trial points performed by the global search algorithm and Hooke--Jeeves 
+method when solving corresponding problems are marked also. 
+The positions of the trial points demonstrate clearly the multiextremality of the function $G(y)$ (the 
+trial points form a nonuniform coverage of the search domain) as well as well as the ravine character 
+of the function $R(x)$ (the trial points form an elongated group along the ravine bottom).
 
-В первой серии экспериментов были решены 100 задач вида (\ref{test_problem}), в которых размерность локальных подзадач равнялась $N=50$.
-Задача считалась решенной, если в процессе минимизации функции $G(y)$ алгоритм глобального поиска сгенерировал точку очередного испытания $y^k$ в $\epsilon$-окрестности глобального минимума $y^*$, т.е.
-$\left\|y^* - y^k\right\|\leq\epsilon$, где $\epsilon = 10^{-2}$.
-При этом точность поиска минимума при минимизации унимодальной функции $R(x)$ выбиралась в 1000 раз меньше, т.е. $\delta = 10^{-5}$.
-В алгоритме глобального поиска использовался параметр $r=3.0$ из (\ref{Rule_R}) и параметр построения развертки $m=10$; использованный локальный метод не имел дополнительных параметров.
+The problem was considered to be solved if in the course of minimization of the function $G(y)$, the 
+global search algorithm generates the next trial point $y^k$ in the $\epsilon$-nearness of the global 
+minimum $y^*$ i. e. $\left\|y^* - y^k\right\|\leq\epsilon$ where $\epsilon = 10^{-2}$. 
+At that, the precision of finding the minimum when minimizing the unimodal function $R(x)$ was 
+selected 1000 times lower i. e. $\delta = 10^{-5}$.
+In the global search algorithm, the parameter $r=4.5$ from (\ref{Rule_R}) and the evolvent building 
+parameter $m=10$ were used; the local method employed has no additional parameters.
 
-%Дать характеристики параллельного запуска - число узлов, процессов, процессоров и т.п.
+In the first series of experiments, 100 problems of the kind (\ref{test_problem}) were solved, in which 
+the dimensionality of the local subproblems was $N=50$. The problems were solved using 1, 4, and 8 
+cluster nodes, in which from 2 up to 32 MPI-processes were employed. According to the 
+parallelization scheme presented in Fig.~\ref{MPI_tree}, the solving of the global optimization 
+problem at the upper nesting level was performed in a single (master) process, which initiated the 
+parallel solving of the local problems of the lower level (from 1 up to 31 problems, respectively). So 
+far, the minimum possible number of processes employed was equals to 2: one per each nesting level.
 
-В table \ref{tab1} отражено среднее число итераций $K_{av}$, выполненных алгоритмом глобального поиска при минимизации функции $G(y)$, среднее время $T_{av}$ решения задачи в целом, а также ускорение по времени $S$  в зависимости от числа $p$ используемых MPI-процессов.
+Table \ref{tab1} presents the averaged number of iterations $K_{av}$ performed by the global search 
+algorithm when minimizing the function $G(y)$, the averaged time of solving the whole problem 
+$T_{av}$, and the speed-up in time $S$ subject to the number of MPI-processes employed $p$.
 
@@
 \centering
-\caption{Результаты решения серии задач вида}\label{tab1}
+\caption{Results of solving a series of problems of dimensionality $N=52$}\label{tab1}
 \begin{tabular}{cccc}
@@
 
-%Привести описание экспериментов на задаче размерности 100+
+In the second series of experiments, 10 problems of the kind (\ref{test_problem}) were solved, in 
+which the dimensionality of the local subproblems was equal $N=100$. The running parameters were 
+the same as in the previous series of experiments. The results of solving are presented in Table 
+\ref{tab2}.
 
-\section{Conclusion} %Заключение надо поправить!!!
+\begin{table}
+\centering
+\caption{Results of solving a series of problems of dimensionality $N=102$}\label{tab2}
+\begin{tabular}{cccc}
+\hline\noalign{\smallskip}
+ $\;\;\;p\;\;\;$  &  $\;\;\;K_{av}\;\;\;$ &  $\;\;\;T_{av}\;\;\;$ & $\;\;\;S\;\;\;$ \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+ 2  & 299.1  & 41.19 & ---  \\
+ 16 & 22.0   & 2.92 & 14.1 \\
+ 32 & 11.7   & 1.50 & 27.4 \\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
 
-Результаты проведенных экспериментов на серии задач оптимизации, в которых можно выделить многоэкстремальную и унимодальную части задачи, показывают, что предложенная в работе схема организации параллельных вычислений обладает хорошим потенциалом параллелизма.
-В частности, при решении серии из 100 задач с общим числом переменных 52 (из которых две переменных оказывали глобальное влияние на целевую функцию, а пятьдесят влияли локально) получено линейное ускорение (в среднем).
+\section{Conclusion}
 
-А при решении задач с числом локально влияющих переменных $N=100$ было получено ускорение ...
+The results of experiments performed on the series of the optimization problems, in which a 
+multiextremal part and a unimodal one can be selected demonstrated the scheme of organization of the 
+parallel computations proposed in the present work to have a good potential for parallelization. 
+In particular, when solving a series of 100 problems with total number of variables equal to 52 (from 
+which, two first variables cause a global effect on the objective function and fifty affected this one 
+locally), a speedup of 24.6 has been obtained when using  32 processes. 
+In the case of solving  the problems with more complex local subproblems (with the number of the 
+local variables $N=100$), the speedup increased up to 27.4 when using 32 processes.
 
-%Проведенное исследование позволяет с уверенностью говорить о решении существенно многомерных задач оптимизации указанного класса.
-
-%Направлением дальнейших исследований может являться как расширение числа используемых локальных и глобальных методов, так и расширения круга рассматриваемых задач, включая задачи оптимизации с ограничениями, а также задачи многокритериальной оптимизации.
+The developed scheme of parallel computations can be applied in the identification of complex 
+mathematical models, which are featured by a large number (tens and hundreds) of unknown 
+parameters.
 

### 2019-11-28 2020/PAVT_Poster/poster-ru.tex

@@
 
-Описание плаката (с текстом в объеме \emph{одной страницы формата А4}) содержит информацию о планах и начальных результатах недавно начатого научного исследования.
-Описание плаката, в отличие от полных и коротких статей, \emph{не имеет аннотации и списка ключевых слов, а также переводов названия, списка авторов и др.
-элементов на английский язык}.
-В то же время описание плаката может содержать рисунки, таблицы и список литературы.
-Страница статьи, представляемой на конференцию, должна иметь \emph{размеры}
-$297 \times 210$~мм. Все \emph{поля} страницы должны иметь одинаковый размер~--- 25~мм. Не допускается использование нумерации, принудительных разрывов страниц и колонтитулов.
+В рамках начатого исследования рассматривается параллельный алгоритм глобальной оптимизации для функций, имеющих одну или несколько точек разрыва типа конечного скачка.
+Разрывы целевой функции могут являться отражением в математической модели специфики оптимизируемого объекта (например, ударными воздействиями, резонансными явлениями, скачками геометричских размеров или свойств материала и т.п.).
+Во многих случаях множество точек разрыва является известным.
+Однако вместе с тем существуют задачи, в которых нет априорных оценок точек разрыва, но известно, что такие точки возможны.
 
-\emph{Название} оформляется шрифтом размером 16~пт с выравниванием по центру страницы. Сверху от остального текста название отделяется одной пустой строкой.
+Задача рассматривается в одномерной постановке
+\[
+\varphi^* = \varphi(x^*)=\min\left\{\varphi(x):x\in[a,b]\right\},
+\] 
+т.к. решение многомерных задач может быть сведено к решению соответствующих им одномерных задач \cite{Grishagin07}.
 
-\emph{Список авторов} оформляется шрифтом размером 12~пт с выравниванием по центру и отделяется от названия одной пустой строкой размером 16~пт.
-Авторы перечисляются через запятую, фамилия ставится справа от инициалов.
-\emph{Организация} оформляется шрифтом размером 12~пт с выравниванием по центру и отделяется от списка авторов одной пустой строкой размером 6~пт.
-В качестве организации необходимо указать полное наименование организации, в которой работают авторы.
-В случае, если авторы из разных организаций, через запятую указывается две или более организаций, и принадлежность автора к соответствующей организации обозначается при помощи сносок.
+Метод распараллен с использованием общего подхода, изложенного в \cite{Strongin13} и успешно примененного для распараллеливания оптимизационных алгоритмов для различных архитектур \cite{Barkalov14,Barkalov16}.
 
-В тексте описания плаката используется \emph{шрифт} Times New Roman.
-\emph{Абзац} оформляется шрифтом размером 11~пт с выравниванием по ширине страницы, одинарным интервалом между строками и автоматической расстановкой переносов.
-Абзацы не разделяются интервалами и начинаются с красной строки с отступом 7~мм.
-Все \emph{рисунки и таблицы} должны иметь подпись c шрифтом размером 10~пт, отступ сверху и снизу 6 пт и выравнивание по центру страницы.
-Подпись к таблице помещается над таблицей, подпись к рисунку~--- под рисунком.
-
-\emph{Благодарности и ссылки на грант} оформляются в виде сноски к названию статьи и обозначаются символом *.
-\emph{Перекрестные ссылки} на литературу заключаются в квадратные скобки и перечисляются через запятую или тире, например \cite{eremin}, \cite{eremin,sokolinsky}, \cite{eremin,sokolinsky,stonebraker}.
-\emph{Библиографические источники} оформляются в виде нумерованного арабскими цифрами списка шрифтом размером 11~пт с выравниванием по левому краю.
+Проведены предварительные эксперименты с распараллеливанием на общей памяти.
+Результаты экспериментов показали ...
+Направлением дальнейших работ будет (i) расширение класса рассматриваемых задач (многомерные задачи с невыпуклыми ограничениями); (ii) распараллеливание алгоритма с использованием распределенной памяти.
 

### 2019-11-29 2020/PAVT_Nested_Full/pavt_paper.tex
Fix some misprints

@@
 
-The global optimization problems arise in various fields of modern science and technology often. For 
+The global optimization problems arise often in various fields of modern science and technology. For 
 example, the problems of finding the configurations of various chemical compounds corresponding to 
@@
 mathematical models based on the data of experiments became a classical one. In the problems of this 
-kind, it is required to carry out a search for the values of the unknown parameters of the model, at 
+type, it is required to carry out a search for the values of the unknown parameters of the model, at 
 which the results of computations according to the model are close to the results obtained 
@@
 objective function is assumed to be multiextremal with respect to a part of variables only. The rest 
-variables have a local effect i. e. the objective function is monoextremal with respect to these ones. 
+variables have a local effect, i.e. the objective function is unimodal with respect to these ones. 
 In this case, the solving of the problem can be organized according to the nested (recursive) 
 optimization scheme. The solving of the multiextremal subproblem, for which the use of complex 
-global search algorithms is required will be performed at the upper level of recursion.
-The monoextremal subproblems (each corresponding to a fixed set of values of the first part of 
-parameters) will be solved at the lower level. For solving the monoextremal problems, the efficient 
+global search algorithms is required, will be performed at the upper level of recursion.
+The unimodal subproblems (each corresponding to a fixed set of values of the first part of 
+parameters) will be solved at the lower level. For solving these subproblems, the efficient 
 methods of local optimization, linear programming, or linear algebra can be applied (subject to the 
@@
 
-The paper text describing the results of the present study, is organized as follows. 
+The paper text describing the results of the present study is organized as follows. 
 In Section 2, the description of the nested optimization scheme taking into account different properties 
@@
 
-Let us consider an optimization problem of the kind
+Let us consider an optimization problem of the type
 \begin{eqnarray}\label{main_problem}
@@
 At the same time, at any fixed set of values $\overline{y}$ the function $f(x,\overline{y})$ is a 
-unimodal one i.e. the function $f(x,\overline{y})$ has a single minimum point $x^*$ (depending, in 
+unimodal one, i.e. the function $f(x,\overline{y})$ has a single minimum point $x^*$ (depending, in 
 general, on $\overline{y}$).
 
-Accounting for such a peculiarity of the considered problem class can reduce the computational costs 
-of the optimum search essentially. Indeed, according to known nested optimization scheme 
+Accounting for such a peculiarity of the considered problem class can reduce essentially the computational costs 
+of the optimum search. Indeed, according to known nested optimization scheme 
 \cite{Carr}, the solving of the initial problem can be reduced to the problem of finding the global 
@@
 \begin{equation}\label{global_problem}
-\varphi(y^*) = \min_{y\in D}  \varphi (y)
+\varphi(y^*) = \min_{y\in D}  \varphi (y),
 \end{equation}
@@
 (\ref{global_problem}) can be considered as a global optimization problem, in which the objective 
-function value can be computed with high precision but this operation is computation-costly.
+function value can be computed with high precision but this operation is time-consuming.
 
 The following approximation problem can be considered as a particular example of a problem, in 
-which different character of dependence on different groups of parameters Takes place. 
+which different character of dependence on different groups of parameters takes place. 
 Let us assume $m$ values $u_j = u(t_j), 1 \leq j \leq m, $ of the function $u(t)$ to have been obtained 
@@
 
-Let us introduce a measure of response of the function $u(t,\omega,c)$ to the experiential data as a 
+Let us introduce a measure of the deviation of the function $u(t,\omega,c)$ from the experiential data as a 
 sum of squares 
@@
 can be done e.g. by Gaussian elimination. At the same time, the external problem (\ref{outer_prob}) 
-will be, generally speaking, a multiextremal one, and its solving will require considerable 
+will be, in general, a multiextremal one, and its solving will require considerable 
 computational resources.
@@
 So far, to solve the problem (\ref{main_problem}) in the nested scheme (\ref{global_problem}), 
-(\ref{local_problem}), one can apply parallel global search algorithm for solving the external 
+(\ref{local_problem}) one can apply parallel global search algorithm for solving the external 
 subproblems (\ref{global_problem}). At that, a set of independent subproblems (\ref{local_problem}) 
@@
 An efficient parallel global optimization algorithm intended for solving the problems of the kind 
-(\ref{global_problem}) has been developed in Lobachevskii University of Nizhnii Novgorod (UNN) 
+(\ref{global_problem}) has been developed in Lobachevsky University of Nizhni Novgorod (UNN) 
 \cite{Barkalov2018,globalizerSystem,Strongin2018}. 
@@
 equal-power sets, and the interval $[0,1]$ can be mapped continuously into $D$ using the Peano 
-curve $y(x)$ i. e. $D = \left\{y(x):0\leq x\leq 1\right\}$.
+curve $y(x)$, i.e. $D = \left\{y(x):0\leq x\leq 1\right\}$.
 
@@
 At the first iteration of the method, the trials are performed in parallel in two boundary points $x^1 = 
-0, \; x^2 = 1$ as well as in $(p-2)$ arbitrary internal points of the interval $[0,1]$ i. e. 
+0, \; x^2 = 1$ as well as in $(p-2)$ arbitrary internal points of the interval $[0,1]$, i.e. 
 $x^i\in(0,1),3\leq i \leq p$.
@@
 of $r$ allows to consider the product $r \mu$ as an estimate
-of the objective function Lipschitz constant $L$.
+of the Lipschitz constant $L$ of the objective function .
 
@@
 the principle of local descent, when the method passes to the points with smaller values of the 
-objective function at every step progressively. Almost all these methods cam be represented in the 
+objective function at every step progressively. Almost all these methods can be represented in the 
 form of the iterative relation 
@@
 The set of the computed local characteristics $I^k=I(x^k)$ may include: the function values $f^k = 
-f( x^k )$, the vector of gradient $\nabla f^k = \nabla f(x^k)$, the matrix of the second derivatives 
+f( x^k )$, the gradient $\nabla f^k = \nabla f(x^k)$, the matrix of the second derivatives 
 (Hessian) $\Gamma_k=\Gamma^f(x^k)$. Particular set of the characteristics to be measured depends 
@@
 In the local optimization problems arising in the applications, the a priori information on the objective 
-functions is quite limited usually. For example, a local character of the dependence of the function on 
+functions is usually quite limited. For example, a local character of the dependence of the function on 
 the parameters may be suggested only whereas the gradient values (and, moreover, Hessian) are
 unknown. In this case, the direct search methods are applied, which don't utilize any suggestions on
-the smoothness of the minimized function. When searching for the minimum, these methods measure 
+the smoothness of the objective function. When searching for the minimum, these methods measure 
 the function values only. The rules of the placement of the iteration points in these ones is based on 
@@
 \begin{equation}\label{test_problem}
-f(x,y) = G(y)+R(x),
+f(x,y) = G(y)+R(x)
 \end{equation}
@@
 \end{eqnarray}
-were considered where
+were considered, where
 \begin{eqnarray} \nonumber
@@
 in the interval $[-1,1]$.
-The class of functions is used for testing the global optimization algorithms often. 
+This class of functions is often used for testing the global optimization algorithms. 
 As the local part of the problem, the modified Rosenbrock function 
@@
 Fig.~\ref{fig_level}(a),(b). These functions are the complex ones for corresponding global and local 
-optimization methods since the functions $G(y)$ are essentially multiextremal ones, and the functions 
-$R(x)$ have a ravine structure manifested clearly.
+optimization methods since the function $G(y)$ is essentially multiextremal one, and the function 
+$R(x)$ has a ravine structure manifested clearly.
 
@@
 The positions of the trial points demonstrate clearly the multiextremality of the function $G(y)$ (the 
-trial points form a nonuniform coverage of the search domain) as well as well as the ravine character 
+trial points form a nonuniform coverage of the search domain) as well as the ravine character 
 of the function $R(x)$ (the trial points form an elongated group along the ravine bottom).
@@
 global search algorithm generates the next trial point $y^k$ in the $\epsilon$-nearness of the global 
-minimum $y^*$ i. e. $\left\|y^* - y^k\right\|\leq\epsilon$ where $\epsilon = 10^{-2}$. 
+minimum $y^*$, i.e. $\left\|y^* - y^k\right\|\leq\epsilon$, where $\epsilon = 10^{-2}$. 
 At that, the precision of finding the minimum when minimizing the unimodal function $R(x)$ was 
-selected 1000 times lower i. e. $\delta = 10^{-5}$.
+selected 1000 times lower, i.e. $\delta = 10^{-5}$.
 In the global search algorithm, the parameter $r=4.5$ from (\ref{Rule_R}) and the evolvent building 
-parameter $m=10$ were used; the local method employed has no additional parameters.
+parameter $m=10$ were used; the local Hooke--Jeeves method employed has no additional parameters.
 
@@
 parallel solving of the local problems of the lower level (from 1 up to 31 problems, respectively). So 
-far, the minimum possible number of processes employed was equals to 2: one per each nesting level.
+far, the minimum possible number of processes employed was equals to 2, one per each nesting level.
 
@@
 algorithm when minimizing the function $G(y)$, the averaged time of solving the whole problem 
-$T_{av}$, and the speed-up in time $S$ subject to the number of MPI-processes employed $p$.
+$T_{av}$, and the speed-up in time $S$ subject to the number of employed MPI-processes $p$.
 
@@
 
-The results of experiments performed on the series of the optimization problems, in which a 
-multiextremal part and a unimodal one can be selected demonstrated the scheme of organization of the 
+The results of experiments performed on the series of the optimization problems
+%, in which a multiextremal part and a unimodal one can be selected,
+demonstrated the scheme of the 
 parallel computations proposed in the present work to have a good potential for parallelization. 
@@
 local variables $N=100$), the speedup increased up to 27.4 when using 32 processes.
-
 The developed scheme of parallel computations can be applied in the identification of complex 

### 2019-11-29 2020/PAVT_Poster/poster-ru.tex
Abstract is ready

@@
 \] 
-т.к. решение многомерных задач может быть сведено к решению соответствующих им одномерных задач \cite{Grishagin07}.
+так как решение многомерных задач может быть сведено к решению соответствующих им одномерных с помощью различных схем редукции размерности \cite{Strongin91}.
 
-Метод распараллен с использованием общего подхода, изложенного в \cite{Strongin13} и успешно примененного для распараллеливания оптимизационных алгоритмов для различных архитектур \cite{Barkalov14,Barkalov16}.
+Последовательный метод решения задач глобальной оптимизации с разрывными функциями (будем ссылаться на него как на Алгоритм Глобального Поиска с Разрывами, АГП-Р) был предложен в \cite{Strongin91}.
+АГП-Р был распараллелен с использованием общего подхода к распараллеливанию алгоритмов глобальной оптимизации, изложенного в \cite{Strongin13} и успешно примененного для различных архитектур \cite{Barkalov14,Barkalov16}.
 
-Проведены предварительные эксперименты с распараллеливанием на общей памяти.
-Результаты экспериментов показали ...
-Направлением дальнейших работ будет (i) расширение класса рассматриваемых задач (многомерные задачи с невыпуклыми ограничениями); (ii) распараллеливание алгоритма с использованием распределенной памяти.
+С помощью параллельного АГП-Р проведена серия экспериментов на тестовых задачах с разрывными функциями.
+Было получено линейное сокращение числа итераций метода в зависимости от числа задействованных потоков, что в предположении большой трудоемкости вычислений значений целевой функции будет соответствовать линейному ускорению по времени.
+Например, при использовании 64 потоков (на одном узле кластера) ускорение по итерациям составляло от 50 до 55 в зависимости от решаемой задачи.
+
+%\begin{table}
+%\centering
+%\caption{Результаты решения тестовой задачи}\label{tab2}
+%\begin{tabular}{cccc}
+%\hline\noalign{\smallskip}
+ %$\;\;\;p\;\;\;$  &  $\;\;\;K_{av}\;\;\;$ &  $\;\;\;T_{av}\;\;\;$ & $\;\;\;S\;\;\;$ \\
+%\noalign{\smallskip}\hline\noalign{\smallskip}
+ %1  & 299.1  & 41.19 & ---  \\
+ %4 & 22.0   & 2.92 & 14.1 \\
+ %16 & 11.7   & 1.50 & 27.4 \\
+%\noalign{\smallskip}\hline
+%\end{tabular}
+%\end{table}
+
+Направлением дальнейших работ будет: (i) расширение класса рассматриваемых задач (многомерные задачи с невыпуклыми ограничениями); (ii) оценка масштабируемости разработанного алгоритма при использовании большого числа ядер/процессоров.
 
@@
 
-\bibitem{Grishagin07}
-Городецкий~С.Ю., Гришагин~В.А. Нелинейное программирование и многоэкстремальная оптимизация. Н.~Новгород: Изд-во ННГУ, 2007. 489~с.
+%\bibitem{Grishagin07} Городецкий~С.Ю., Гришагин~В.А. Нелинейное программирование и многоэкстремальная оптимизация. Н.~Новгород: Изд-во ННГУ, 2007. 489~с.
 

### 2019-11-30 2020/PAVT_Sovrasov/pavt_paper.tex
Add a draft of the next PAVT paper

\documentclass[11pt, oneside, a4paper]{article}
%\usepackage[cp1251]{inputenc} % кодировка
\usepackage[utf8]{inputenc} % кодировка
\usepackage[english, russian]{babel} % Русские и английские переносы
\usepackage{graphicx}          % для включения графических изображений
\usepackage{cite}              % для корректного оформления литературы
\usepackage{enumitem}
\usepackage{pavt-ru}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{cite}
%\usepackage{subfig}
%\usepackage{caption}
%\captionsetup[table]{skip=9pt}
%\captionsetup{labelfont=bf}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\sign}{sign}
\newtheorem{theorem}{Теорема}

\begin{document}

% \title - название статьи
% \authors - список авторов

\title{Параллельный алгоритм для получения равномерного приближения решений множества задач глобальной оптимизации с нелинейными ограничениями
\footnote{Исследование выполнено при поддержке РНФ, проект №\,16-11-10150.}}

\authors{В.В.~Соврасов}
\organizations{Нижегородский государственный университет им. Н.И. Лобачевского}

% Аннотация заключается в окружение abstract
\begin{abstract}
В данной работе рассматривается построение параллельной версии алгоритма глобальной оптимизации, решающего одновременно множество задач с нелинейными ограничениями и получающего при этом равномерные оценки решений на этом множестве. Последнее свойство позволяет наиболее оптимально распределять вычислительные ресурсы, т.к. в процессе работы алгоритма погрешности численного решения во всех задачах убывают примерно с одинаковой скоростью. Подобные серии задач возникают, если задача глобальной оптимизации имеет дискретный параметр или при решении задачи многокритериальной оптимизации методом свётртки критериев. Рассматриваемый алгоритм использует отображения типа кривой Пеано для редукции многомерных задач оптимизации к одномерным. Эффективность реализованного алгоритма протестирована на наборах искусственно сгенерированных задач глобальной оптимизации.
\end{abstract}

\keywords{Computational mathematics / Вычислительная математика, Parallel and distributed computing technologies / Технологии параллельных и распределенных вычислений}

% \section{название} - заголовок раздела первого уровня
% \subsection{название} - заголовок раздела второго уровня
% \subsubsection{название} - заголовок раздела третьего уровня
% Не используйте уровень вложенности заголовков больше трех!
% Каждый абзац текста в статье начинается командой \par или пустой
% строкой.

\section{Введение}

Нелинейная глобальная оптимизация невыпуклых функций традиционно считается одной из самых трудных
задач математического программирования. Отыскание глобального минимума функции от нескольких переменных
зачастую оказывается сложнее, чем локальная оптимизация в тысячемерном пространстве. Для последней может оказаться достаточно
применения простейшего метода градиентного спуска, в то время как чтобы \textit{гаранитрованно} отыскать глобальный оптимум методам
оптимизации приходится накапливать информацию о поведении целевой функции во всей области поиска \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}. Решение серии таких задач при ограниченных вычислительных
ресурсах является ещё более сложной проблемой: помимо поиска глобального экстремума необходимо
распределять вычислительные ресурсы так, чтобы сразу во всех решаемых задачах положение глобального
экстремума было оценено примерно с одинаковым качеством. Обычно серию из \(q\) задач решают либо последовательно, либо
параллельно порциями по \(p,\:p<<q\) задач, где \(p\) --- количество параллельных вычислительных усройств.
Такой подход ведёт к тому, что в каждый момент времени до окончания вычислений
остаются задачи, в которых оценка глобального оптимума не получена вообще, в то время, как в задачах из начала
списка оптимум может быть оценён даже с избыточной точностью.

В данной работе рассматривается обобщение ранее разработанного в ННГУ им. Н. И. Лобачевского
параллельного метода глобальной оптимизации для одновременного решения множества задач \cite{BarkalovStrongin2018} на
случай задач с нелинейными ограничениями. Для учёта ограничений используется индексная схема \cite{Strongin2000},
позволяющая работать с частично вычислимыми целевым функциями и обладающая экономичностью,
сравнимой с другими подходами \cite{BarkalovLebedev2017}. Эффективность реализованного
алгоритма показана на примере решения множеств задач, сгенерированных специализированным
механизмом, порождающим наборы задач заданной размерности с заданным количеством нелинейных ограничений \cite{GergelBarkalov2019}.
Кроме искуственно сгенерированных задач, рассматриваемый метод протестирован также
на множестве задач, возникающем при решении задачи многокритериальной оптимизации
с нелинейными ограничениями методом свёртки критериев \cite{Ehrgott2005}.

\section{Постановка задачи глобальной оптимизации}
В рамках данной работы будем рассмативать следующую постановку задачи глобальной
оптимизации: найти глобальный минимум \(N\)-мерной функции \(\varphi(y)\) в гиперинтервале
\(D=\{y\in \mathbf{R}^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
Для построения оценки глобального минимума по конечному количеству вычислений
значения функции требуется, чтобы скорость изменения \(\varphi(y)\) в \(D\) была ограничена.
В качестве такого ограничения как правило принимается условие Липшица.
\begin{displaymath}
\label{task}
\varphi(y^*)=\min\{\varphi(y):y\in D\}
\end{displaymath}
\begin{displaymath}
\label{lip}
|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
\end{displaymath}

Существуют различные методы, решающие рассмотренную многомерную задачу напрямую \cite{SergeyevKvasov2017, Jones2009},
а также эффективные методы решения одномерных задач \cite{Norkin1992, Strongin2000}. В данной работе рассматривается одномерный метод,
который применяется совместно со схемой редукции размерности.
Классической схемой редукции размерности исходной задачи для алгоритмов глобальной оптимизации является
использование разверток --- кривых, заполняющих пространство \cite{Sergeyev2013}.
\begin{equation}
\label{cube}
\lbrace y\in \mathbf{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
\end{equation}

Отображение вида (\ref{cube}) позволяет свести задачу в многомерном пространстве к решению
одномерной ценой ухудшения её свойств. В частности, одномерная функция \(\varphi(y(x))\)
является не Липшицевой, а Гёльдеровой:
\begin{displaymath}
\label{holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0;1],
\end{displaymath}
где константа Гельдера \(H\) связана с константой Липшица \(L\) соотношением
\begin{displaymath}
H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}.
\end{displaymath}

Область \(D\) также может быть задана с помощью функциональных ограничений, что
значительно усложняет задачу.
Постановка задачи глобальной оптимизации в этом случае будет иметь следующий вид:
\begin{equation}
  \label{eq:constrained_problem}
  \varphi(y^*)=\min\{\varphi(y):g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}
\end{equation}
Обозначим \(g_{m+1}(y)=\varphi(y)\). Далее будем предполагать, что все функции \(g_k(y),1\leqslant k \leqslant m+1\)
удовлетворяют условию Липшица в некотором гиперинтервале, включающем \(D\).

Далее будем интересоваться решением серии из \(q\) задач вида (\ref{eq:constrained_problem}):
\begin{equation}
  \label{eq:many_problems}
  \min\left\{\varphi_1(y), y\in D_1 \right\}, \min\left\{\varphi_2(y), y\in D_2\right\},..., \min\left\{\varphi_q(y), y\in D_q\right\}.
\end{equation}

\section{Описание метода глобальной оптимизации}

Принимая во внимание схему редукции размерности (\ref{cube}), будем при описании метода считать, что
требуется найти глобальный минимум функции \(\varphi(x), x\in[0;1]\),
удовлетворяющей условию Гёльдера, при ограничениях \(g_j(x)\), также
удовлетворяющих этому условию на интервале \([0;1]\).

Рассматриваемый индексный алгоритм глобального поиска (ИАГП) для решения
одномерной задачи (\ref{eq:constrained_problem}) предполагает построение последовательности
точек \(x_k\), в которых вычисляются значения минимизируемой функции или ограничений \(z_k = g_s(x_k)\).
Для учёта последних используется индексная схема \cite{Strongin2000}. Пусть \(Q_0=[0;1]\). Ограничение, имеющее номер
 \(j\), выполняется во всех точках области
\begin{displaymath}
  Q_j=\left\{x\in [0;1]:g_j(x)\leq 0\right\},
\end{displaymath}
которая называется допустимой для этого ограничения. При этом допустимая область \(D\)
исходной задачи определяется равенством: \(D=\cap _{j=0}^{m}Q_{j}\).
Испытание в точке \(x\in [0;1]\) состоит в последовательном вычислении значений
величин \(g_{1}(x),...,g_{\nu }(x)\), где значение индекса \(\nu\) определяется условиями:
\(x\in Q_{j},0\leqslant j<\nu ,x\notin Q_{\nu }\). Выявление первого нарушенного ограничения
прерывает испытание в точке \(x\). В случае, когда точка \(x\)  допустима, т. е.
\(x\in D\) испытание включает в себя вычисление всех функций задачи. При этом значение
индекса принимается равным величине \(\nu =m+1\). Пара \(\nu =\nu (x),z=g_{\nu }(x)\),
где индекс \(\nu\) лежит в границах \(1\leqslant \nu \leqslant m+1\), называется результатом
испытания в точке \(x\).

Такой подход к проведению испытаний позволяет свести исходную задачу с функциональными
ограничениями к безусловной задаче минимизации разрывной функции:

\begin{displaymath}
  \begin{array}{lr}
    \psi (x^{*})=\min_{x\in [0;1]}\psi (x), \\
    \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu }&\nu <M\\(g_{M}(x)-g_{M}^{*})/H_{M}&\nu =M\end{cases}}
  \end{array}
\end{displaymath}

Здесь \(M=\max_{}^{}\left\{\nu (x):x\in [0;1]\right\}\), а \(g_{M}^{*}=\min _{}^{}\left\{g_{M}(x):x\in \cap _{i=0}^{M-1}Q_{i}\right\}\).
В силу определения числа \(M\), задача отыскания \(g_{M}^{*}\)
всегда имеет решение, а если \(M=m+1\), то \(g_{M}^{*}=\varphi(x^{*})\).
Дуги функции \(\psi (x)\) гельдеровы на множествах \(\cap _{i=0}^{j}Q_{i},0\leq j\leq M-1\)
с константой 1, а сама \(\psi (x)\) может иметь разрывы первого рода на границах этих множеств.
Несмотря на то, что значения констант Гёльдера \(H_k\) и величина \(g_{M}^{*}\) заранее неизвестны,
они могут быть оценены в процессе решения задачи.

Множество троек \(\{(x_k,\nu_k,z_k)\}, 1\leqslant k\leqslant n\) составляет поисковую информацию,
накопленную методом после проведения \(n\) шагов.

На первой итерации метода испытание проводится в произвольной внутренней точке \(x_1\)
интервала \([0;1]\). Индексы точек 0 и 1 считаются нулевыми, значения \(z\) в
них не определены. Пусть выполнено \(k\geqslant 1\) итераций метода,
в процессе которых были проведены испытания в \(k\) точках \(x_i, 1\leqslant i\leqslant k\).
Тогда точкa \(x^{k+1}\) поисковых испытаний следующей \((k+1)\)-ой
итерации определяются в соответствии с правилами:

Шаг 1. Перенумеровать точки множества \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\),
которое включает в себя граничные точки интервала \([0;1]\), а также точки предшествующих
испытаний, нижними индексами в порядке увеличения значений координаты, т.е.
\begin{displaymath}
0=x_0<x_1<\dotsc<x_{k+1}=1
\end{displaymath}
и сопоставить им значения \(z_{i}=g_{\nu }(x_{i}),\nu =\nu (x_{i}),i={\overline {1,k}}\).

Шаг 2. Для каждого целого числа \(\nu ,1\leqslant \nu \leqslant m+1\) определить соответствующее
ему множество \(I_{\nu }\) нижних индексов точек, в которых вычислялись значения
функций \(g_{\nu }(x)\):
\begin{displaymath}
  I_{\nu }=\{i:\nu (x_{i})=\nu ,1\leqslant i\leqslant k\},1\leq \nu \leqslant m+1,
\end{displaymath}
определить максимальное значение индекса \(M=\max\{\nu (x_{i}),1\leq i\leq k\}\).

Шаг 3. Вычислить текущие оценки для неизвестных констант Гёльдера:
\begin{equation}
  \label{step2}
  \mu _{\nu }=\max\{\frac{|g_{\nu }(x_{i})-g_{\nu }(x_{j})|}{(x_{i}-x_{j})^{\frac{1}{N}}}:i,j\in I_{\nu },i>j\}.
\end{equation}
Если множество \(I_{\nu }\) содержит менее двух элементов или если значение \(\mu _{\nu }\)
оказывается равным нулю, то принять \(\mu _{\nu }=1\).

Шаг 4. Для всех непустых множеств \(I_{\nu },\nu ={\overline {1,M}}\) вычислить оценки
\begin{displaymath}
  z_{\nu }^{*}={\begin{cases}\min\{g_{\nu }(x_{i}):x_{i}\in I_{\nu }\}&\nu =M\\-\varepsilon _{\nu }&\nu <M\end{cases}},
\end{displaymath}
где вектор с неотрицательными координатами \(\varepsilon _{R}=(\varepsilon _{1},..,\varepsilon _{m})\) называется вектором резервов.

Шаг 5. Для каждого интервала \((x_{i-1};x_{i}),1\leqslant i\leqslant k\) вычислить характеристику
\begin{equation}
  \label{step3_1}
  R(i)={\begin{cases}\Delta _{i}+{\frac {(z_{i}-z_{i-1})^{2}}{(r_{\nu }\mu _{\nu })^{2}\Delta _{i}}}-2{\frac {z_{i}+z_{i-1}-2z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu (x_{i})=\nu (x_{i-1})\\2\Delta _{i}-4{\frac {z_{i-1}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu (x_{i-1})>\nu (x_{i})\\2\Delta _{i}-4{\frac {z_{i}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu (x_{i})>\nu (x_{i-1})\end{cases}}
\end{equation}
где \(\Delta _{i}=(x_{i}-x_{i-1})^{\frac{1}{N}}\). Величины \(r_{\nu }>1,\nu ={\overline {1,m}}\)
являются параметрами алгоритма. От них зависят произведения \(r_{\nu }\mu _{\nu }\),
используемые при вычислении характеристик в качестве оценок неизвестных констант Гёльдера.

Шаг 6. Выбрать наибольшую характеристику:
\begin{equation}
\label{step4}
t=\argmax_{1\leqslant i \leqslant k+1}R(i)
\end{equation}

Шаг 7. Провести очередное испытание в середине интервала \((x_{t-1};x_{t})\),
если индексы его концевых точек не совпадают: \(x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})\).
В противном случае провести испытание в точке
\begin{displaymath}
  x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})-\operatorname {sgn}(z_{t}-z_{t-1}){\frac {|z_{t}-z_{t-1}|^{n}}{2r_{\nu }\mu _{\nu }^{n}}},\nu =\nu (x_{t})=\nu (x_{t-1}),
\end{displaymath}
а затем увеличить \(k\) на 1.

Алгоритм прекращает работу, если выполняется условие \(\Delta_{t}\leqslant \varepsilon\),
где \(\varepsilon>0\) есть заданная точность. В качестве оценки глобально-оптимального решения выбираются значения
\begin{equation}
\varphi_k^*=\min_{1\leqslant i \leqslant k}\varphi(x_i), x_k^*=\argmin_{1\leqslant i \leqslant k}\varphi(x_i)
\end{equation}

Далее следуя подходу, описанному в \cite{BarkalovStrongin2018}, для решения серии задач (\ref{eq:many_problems}) будем
использовать \(q\) синхронно работающих копий ИАГП с тем лишь отличием, что на шаге 6 при выборе
интервала с наилучшей характеристикой, выбор будет осуществляться из всех интервалов, которые
породили на данный момент \(q\) копий ИАГП. Если наибольшая характеристика соответствует
задаче \(i\), то выполняется шаг 7 в копии метода с номером \(i\), а остальные копии метода простаивают.
Таким образом, на каждой итерации испытание проводится в задаче, наиболее перспективной с точки зрения
характеристик (\ref{step3_1}), что позволяет динамически распределять ресурсы метода между задачами.

В \cite{BarkalovStrongin2018} приведена теория сходимости такого подхода на случай решения задач без ограничений.
При наличии ограничений характеристики интервалов, на концах которых нарушено разное количество ограничений,
вычисляются в соответствии с нижними строчками из (\ref{step3_1}). Нетрудно заметить, что и в этом случае
величины характеристик нормированы, а в случае точных оценок \(z_{\nu }^{*}\) и \(\mu _{\nu }\) их
масштаб не зависит от целевой функции и ограничений конкретной задачи. Таким образом, рассуждения
из \cite{BarkalovStrongin2018} можно провести и в случае использования индексной схемы.

Параллельная модификация метода не отличается от рассматриаемой в \cite{BarkalovStrongin2018}
и заключается в выборе \(p\) интервалов на шаге 6 и выполнения \(p\) испытаний параллельно
на следующем шаге. При этом все ресурсы метода в рамках итерации могут быть направлены как на одну, так и
на \(l\leqslant p\) задач одновременно (в зависимости от того, какой из задач принадлежат выбранные методам интервалы).

\section{Результаты численных экспериментов}

Использование сгенерированных некоторыми случайными механизмами
наборов тестовых задач с известными решениями является одним из общепринятых подходов
к сравнению алгоритмов оптимизации \cite{Beiranvand2017}. В данной работе
будем использовать два генератора тестовых задач, порождающих задачи различной природы \cite{grishaginClass, Gaviano2003}.
Эти генераторы порождают задачи без нелинейных ограничений, поэтому в дополнение к ним использована
система GCGen \footnote{Исходный код системы доступен по ссылке https://github.com/UNN-ITMM-Software/GCGen} \cite{GergelBarkalov2019}, позволяющая генерировать задачи с ограничениями на основе произвольных нелинейных
функций.

Вместе с системой GCGen распростаняются примеры её использования и построения
наборов задач, каждая из которых состоит из целевой функции и двух ограничений,
порождённых генератором \(F_{GR}\) \cite{grishaginClass} или GKLS \cite{Gaviano2003}.

Обозначим набор из 100 задач, полученных с помощью GCGen и первого генератора из \cite{grishaginClass} как \(F_{GR}^C\). Механизм построения функций \(F_{GR}\) не предусматривает контроля за размерностью (она равна 2) и количеством локальных оптимумов, однако известно, что порождаемые функции
являются существенно многоэкстремальными.

Генератор GKLS \cite{Gaviano2003} позволяет получать функции заданной размерности и с заданным количеством экстремумов.
В сочетании с GCGen было порождено три класса по 100 задач размерностей 2, 3, 4. Каждая из задач имеет два ограничения.

Будем считать, что тестовая задача решена, если метод оптимизации провёл очередное испытание \(y^k\) в
\(\delta\)-окрестности глобального минимума \(y^*\), т.е. $\left\|y^k-
y^*\right\|\leqslant \delta = 0.01\left\|b-a\right\|$, где \(a\) и \(b\) --- левая и правая границы гиперкуба из (\ref{eq:task}).
Если указанное соотношение не выполнено до истечения лимита на количество испытаний, то задача считается нерешённой.
Максимальный лимит на количество испытаний установлен для каждого класса задач, в соответствии с размерностью и сложностью (см. таблицу \ref{tab:limits}).

В качестве характеристик метода оптимизации на каждом из классов будем рассматривать среднее число
испытаний, затраченное для решения одной задачи, и количество решённых задач. Чем меньше число испытаний, тем быстрее метод сходится
к решению, а значит и меньше обращается к потенциально трудоёмкой процедуре вычислений целевой функции.
Количество решённых задач говорит о надёжности метода с заданными параметрами на решаемом классе тестовых задач.
Чтобы сделать величины, характеризующие надёжность и скорость сходимости, при подсчёте среднего количества испытаний учитывались только решённые задачи.

Реализация параллельного метода была выполнена на языке C++ с использованием технологии OpenMP
для распареллеливания процесса проведения испытаний на общей памяти. Все вычислительные
эксперименты проведены на машине со следующей конфигурацией: Intel Core i7, 32GB RAM, ОС Unubtu 16.04.

\subsection{Результаты решения сгенерированных задач}

\subsection{Пример решения многокритериальной задачи}

Для демонстрации эффективности подхода к балансировке нагрузки рассмотрим пример,
в котором множество задач вида (\ref{eq:many_problems}) порождено в результате скаляризации
многокритериальной задачи оптимизации с ограничениями.

Рассмотрим тестовую задачу, предложенную в \cite{BinhKorn1999}:
\begin{equation}
  \label{eq:mco_probem}
  \begin{array}{l}
      Minimize \left \{
      \begin{array}{l}
        f_1(y) = 4 y_1^2 + 4 y_2^2 \\
        f_2(y) = (y_1-5)^2 + (y_2-5)^2 \\
      \end{array}
      \right .
      y_1\in [-1;2],y_2\in [-2;1]
      \\s.t.
      \\
      \left \{
      \begin{array}{l}
        g_1(y) = (y_1 - 5)^2 + y_2^2 - 25 \leqslant 0 \\
        g_2(y) = -(y_1 - 8)^2 + -(y_2 + 3)^2 + 7.7 \leqslant 0\\
      \end{array}
      \right .
  \end{array}
\end{equation}

Будем использовать свёртку Гермейера для скаляризации задачи (\ref{eq:mco_probem}).
После свёртки скалярная целевая функция имеет вид:
\begin{equation}
  \varphi(y,\lambda_1,\lambda_2)=\max\{\lambda_1 f_1(y), \lambda_2 f_2(y)\},
\end{equation}
где \(\lambda_1,\lambda_2\in[0,1],\: \lambda_1+\lambda_2=1\). Перебирая все возможные
коэффициенты свёртки, можно найти всё множество парето-оптимальных решений в
задаче (\ref{eq:mco_probem}). Для численного построения множества Парето выберем
100 наборов коэффициентов \((\lambda_1,\lambda_2)\) таких, что
\(\lambda_1^i=i h,\: \lambda_2^i=1-\lambda_1^i,\: h=10^{-2},i=\overline{1, 100}\).

\section{Заключение}

\bibliographystyle{spmpsci}
\bibliography{bibliography}{}
%\begin{biblio}
%\bibitem{eremin}
%Ерёмин~И.И. Фейеровские методы для задач выпуклой и линейной оптимизации.
%Челябинск: Изд-во ЮУрГУ, 2009. 200~с.
%\bibitem{levin}
%Левин~В.К. Отечественные суперкомпьютеры семейства МВС. URL:~\href{http://parallel.ru/mvs/levin.htm}{http://parallel.ru/mvs/levin.htm} (дата обращения: 27.05.2012).
%\end{biblio}

\end{document}

### 2019-12-01 2020/PAVT_Poster/poster-ru.tex
Fix some misprints

@@
 В рамках начатого исследования рассматривается параллельный алгоритм глобальной оптимизации для функций, имеющих одну или несколько точек разрыва типа конечного скачка.
-Разрывы целевой функции могут являться отражением в математической модели специфики оптимизируемого объекта (например, ударными воздействиями, резонансными явлениями, скачками геометричских размеров или свойств материала и т.п.).
+Разрывы целевой функции могут являться отражением в математической модели специфики оптимизируемого объекта (например, ударных воздействий, резонансных явлений, скачков геометрических размеров или свойств материала и т.п.).
 Во многих случаях множество точек разрыва является известным.
@@
 Например, при использовании 64 потоков (на одном узле кластера) ускорение по итерациям составляло от 50 до 55 в зависимости от решаемой задачи.
-
-%\begin{table}
-%\centering
-%\caption{Результаты решения тестовой задачи}\label{tab2}
-%\begin{tabular}{cccc}
-%\hline\noalign{\smallskip}
- %$\;\;\;p\;\;\;$  &  $\;\;\;K_{av}\;\;\;$ &  $\;\;\;T_{av}\;\;\;$ & $\;\;\;S\;\;\;$ \\
-%\noalign{\smallskip}\hline\noalign{\smallskip}
- %1  & 299.1  & 41.19 & ---  \\
- %4 & 22.0   & 2.92 & 14.1 \\
- %16 & 11.7   & 1.50 & 27.4 \\
-%\noalign{\smallskip}\hline
-%\end{tabular}
-%\end{table}
 

### 2019-12-01 2020/PAVT_Sovrasov/pavt_paper.tex
PAVT: update paper

@@
 \usepackage{cite}
-%\usepackage{subfig}
-%\usepackage{caption}
+\usepackage{subfig}
+\usepackage{caption}
 %\captionsetup[table]{skip=9pt}
@@
 В качестве такого ограничения как правило принимается условие Липшица.
-\begin{displaymath}
-\label{task}
+\begin{equation}
+\label{eq:task}
 \varphi(y^*)=\min\{\varphi(y):y\in D\}
-\end{displaymath}
+\end{equation}
 \begin{displaymath}
@@
 
-Обозначим набор из 100 задач, полученных с помощью GCGen и первого генератора из \cite{grishaginClass} как \(F_{GR}^C\).
+% Обозначим набор из 50 задач, полученных с помощью GCGen и первого генератора из \cite{grishaginClass} как \(F_{GR}^C\).
 Механизм построения функций \(F_{GR}\) не предусматривает контроля за размерностью (она равна 2) и количеством локальных оптимумов, однако известно, что порождаемые функции
-являются существенно многоэкстремальными.
+% являются существенно многоэкстремальными.
 
 Генератор GKLS \cite{Gaviano2003} позволяет получать функции заданной размерности и с заданным количеством экстремумов.
-В сочетании с GCGen было порождено три класса по 100 задач размерностей 2, 3, 4. Каждая из задач имеет два ограничения.
+В сочетании с GCGen было порождено 100 задач размерности 2. Каждая из задач имеет два ограничения.
+Также с целью демонстарции того, что свойства метода сохраняются при существенно разных свойствах задач
+был сгенерирован смешанный класс, состоящий из 50 задач с функциями GKLS и 50 задач с функциями \(F_{GR}\).
+На рис. \ref{fig:isolines} предствалены примеры линий уровня рассматриваемых задач. Допустимая область отмечена зелёным цветом.
+
+\begin{figure}[ht]
+    \centering
+    \subfloat[]{
+    {\includegraphics[width=.5\textwidth]{images/2.png}}}
+    \subfloat[]{
+    {\includegraphics[width=.5\textwidth]{images/4.png}}}
+    \caption{Линии уровня и точки испытаний ИАГП в двух синтетических задачах.}
+    \label{fig:isolines}
+\end{figure}
 
@@
 Если указанное соотношение не выполнено до истечения лимита на количество испытаний, то задача считается нерешённой.
-Максимальный лимит на количество испытаний установлен для каждого класса задач, в соответствии с размерностью и сложностью (см. таблицу \ref{tab:limits}).
 
-В качестве характеристик метода оптимизации на каждом из классов будем рассматривать среднее число
-испытаний, затраченное для решения одной задачи, и количество решённых задач. Чем меньше число испытаний, тем быстрее метод сходится
-к решению, а значит и меньше обращается к потенциально трудоёмкой процедуре вычислений целевой функции.
-Количество решённых задач говорит о надёжности метода с заданными параметрами на решаемом классе тестовых задач.
-Чтобы сделать величины, характеризующие надёжность и скорость сходимости, при подсчёте среднего количества испытаний учитывались только решённые задачи.
+%В качестве характеристик метода оптимизации на каждом из классов будем рассматривать среднее число
+%испытаний, затраченное для решения одной задачи, и количество решённых задач. Чем меньше число испытаний, тем быстрее метод сходится
+%к решению, а значит и меньше обращается к потенциально трудоёмкой процедуре вычислений целевой функции.
+При оценке качества метода и его реализации кроме ускорения от распараллеливания и времени выполнения также будем принимать во внимание среднее максимальное расстояния (в смысле \(l_{\inf}\)-нормы) текущей оценки оптимума до его реального положения,
+вычисленное на множестве задач (\ref{eq:many_problems}): \(D_{avg}\) и \(D_{max}\). Динамика этих величи в процессе оптимизации
+показывает, насколько равномерно метод распределяет ресурсы между задачами.
 
@@
 для распареллеливания процесса проведения испытаний на общей памяти. Все вычислительные
-эксперименты проведены на машине со следующей конфигурацией: Intel Core i7, 32GB RAM, ОС Unubtu 16.04.
+эксперименты проведены на машине со следующей конфигурацией: Intel Core i7-7800X, 64GB RAM, Unubtu 16.04 ОS, GCC 5.5 compiler.
 
 \subsection{Результаты решения сгенерированных задач}
+
+Результаты решения тестовых задач последовательной и параллельной версией модифицированного ИАГП
+для решения множества задач представлены в таблице \ref{tab:speedup}. Для всех классов задач параметр \(r=4.7\).
+Во всех экспериментах в целевые функции и ограничения была внесена дополнительная вычислительная нагрузка так,
+чтобы время одного обращения к функции задачи было равно примерно 1мс.
+
+Из таблицы видно, что ускорение по итерациям \(S_i\) растёт линейно с увеличением числа потоков \(p\),
+в то время, как укорение по времени \(S_p\) увеличивается не так быстро, что говорит о неидеальной
+реализации алгоритма. Увеличить реальное ускорение, верхней границей для которого является
+\(S_i\) возможно путём оптимизации взаимодействий между копиями ИАГП и это планируется сделать в ходе будущей работы.
+
+\begin{table}
+  \centering
+  \caption{Результаты экспериментов на наборах синтетических задач}
+  \label{tab:speedup}
+  \begin{tabular}{c|c|cccc}
+    %\cline{1-8}\noalign{\smallskip}
+    Класс задач & \textit{p} & Количество итераций & Время, с & \(S_i\) & \(S_t\)   \\
+    %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
+    \hline
+    GKLS-based \
+         & 1 & 59066 & 97.53 & -    & - \\
+         & 2 & 29060 & 60.56 & 2.04 & 1.61 \\
+         & 4 & 14266 & 38.92 & 4.14 & 2.51 \\
+         & 6 & 9436  & 29.53 & 6.26 & 3.30 \\
+    \hline
+    GKLS \& \(F_{GR}\) based \
+         & 1 & 51434 & 90.20 & -    & - \\
+         & 2 & 25698 & 56.96 & 2.00 & 1.58 \\
+         & 4 & 13015 & 36.67 & 3.95 & 2.46 \\
+         & 6 & 8332  & 26.85 & 6.17 & 3.36 \\
+        \noalign{\smallskip}\hline
+  \end{tabular}
+\end{table}
+
+Для того, чтобы показать равномерную сходимость все тестовые задачи были также решены ИАГП
+в режиме решения отдельных задач. На рис. \ref{fig:devs_mixed} указаны графики величин
+средних и максимальных расстояний от реальных оптимумов до их текущих оценок при решении
+серии из задач, порождённых двумя разными генераторами, по отдельности (сплошная кривая) и совместно (пунктирная кривая).
+Не смотря на значительную разницу в структуре задач, модифицированный ИАГП
+значительно быстрее уменьшает как максимальное, так и среднее отклонение оценок от оптимумов.
+Это говорит о наличии равномерной сходимости по всему множеству совместно решаемых задач.
+При этом в случае последовательного решения задач величина \(D_{max}\) имеет наибольшее значение вплоть
+до решения последней задачи.
+
+\begin{figure}[ht]
+    \centering
+    \subfloat[]{
+    \label{fig:max_dev} {\includegraphics[width=.5\textwidth]{images/mixed_2d_max.pdf}}}
+    \subfloat[]{
+    \label{fig:avg_dev} {\includegraphics[width=.5\textwidth]{images/mixed_2d_avg.pdf}}}
+    \caption{Динамика величин \(D_{avg}\) и \(D_{max}\) в процессе решения множества задач}
+    \label{fig:devs_mixed}
+\end{figure}
 
@@
         g_1(y) = (y_1 - 5)^2 + y_2^2 - 25 \leqslant 0 \\
-        g_2(y) = -(y_1 - 8)^2 + -(y_2 + 3)^2 + 7.7 \leqslant 0\\
+        g_2(y) = -(y_1 - 8)^2 - (y_2 + 3)^2 + 7.7 \leqslant 0\\
       \end{array}
@@
 
+В качестве ограничения на вычислительные ресурсы был выбран лимит в 2500 испытаний.
+Множество вспомогательных скалярных задач решалось двумя способами:
+\begin{itemize}
+  \item каждая задача решается отдельно с помощью ИАГП с установленным лимитом в
+  25 испытаний. Таким образом, вычислительные ресурсы равномерно распределены между задачами;
+  \item все задачи решаются одновременно с помощью обобщённого ИАГП с установленным лимитом в
+  2500 испытаний.
+\end{itemize}
+В обоих случаях параметр \(r=4\).
+
+\begin{figure}[ht]
+    \centering
+    \subfloat[ИАГП, раздеьное решение задач]{
+    \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{images/single_mco.pdf}}}
+    \subfloat[ИАГП для множества задач]{
+    \label{fig:mco_pareto_2} {\includegraphics[width=.5\textwidth]{images/multi_mco.pdf}}}
+    \caption{Численные оценки множества Парето в задаче (\ref{eq:mco_probem}), полученные после 2500 испытаний}
+    \label{fig:mco_pareto}
+\end{figure}
+
+На рис. \ref{fig:mco_pareto} представлены графики решений, полученных каждым из методов.
+Все графики качественно совпадают с указанным в \cite{BinhKorn1999} (авторы не предоставили другой
+информации и решениях для сравнения). Видно, что на рис. \ref{fig:mco_pareto_1}
+кривая Парето имеет вогнутости, что не соответствует решению, указанному в \cite{BinhKorn1999} и
+означает нехватку ресурсов для решении вспомогательных задач.
+Для оценки качества решения также был вычислен показатель \(Spacing(SP)\) \cite{RiquelmeLucken2015},
+характеризующий плотность точек аппроксимации множества Парето.
+\begin{displaymath}
+  SP(S)=\sqrt{\frac{1}{|S|-1} \sum_{i=1}^{|S|} (\overline{d}-d_i)},
+  \:\overline{d}=mean\{d_i\},\:d_i=\min_{s_i,s_j\in S:s_i\ne s_j}||F(s_i)-F(s_j)||_1,\: F=(f_1,f_2)
+\end{displaymath}
+В случае отдельного решения задач \(SP_{single}=0.984\), а при решении задач методом с балансировкой нагрузки
+\(SP_{multi}=0.749\), что говорит о более качественном приближении решения.
+
 \section{Заключение}
+
+В ходе работы была реализована поддержка нелинейных ограничений в алгоритме, решающим
+множество задач глобальной оптимизации. Проведены численные эксперименты, демонстрирующие
+преимущество рассматриваемого подхода над решением задач по отдельности. Показана эффективность
+совместного решения множества задач на примере решения многокритериальной задачи с
+нелинейными ограничениями.
+
+В ходе дальнейшей работы планируется улучшить текущую реализацию алгоритма,
+сократив расходы на содержание поисковой информации для множества задач и тем самым улучшив
+показатели параллельного ускорения по времени. Также планируется реализовать версию
+рассматриваемого алгоритма, работающего на распределённой памяти по схеме, описанной в \cite{BarkalovLebedev2017}.
 

### 2019-12-02 2020/PAVT_Sovrasov/pavt_paper.tex
PAVT: fix misprints

@@
 В данной работе рассматривается построение параллельной версии алгоритма глобальной оптимизации, решающего одновременно множество задач с нелинейными ограничениями и получающего при этом равномерные оценки решений на этом множестве.
-Последнее свойство позволяет наиболее оптимально распределять вычислительные ресурсы, т.к.
+Последнее свойство позволяет наиболее оптимально распределять вычислительные ресурсы, т.
+к.
 в процессе работы алгоритма погрешности численного решения во всех задачах убывают примерно с одинаковой скоростью.
-Подобные серии задач возникают, если задача глобальной оптимизации имеет дискретный параметр или при решении задачи многокритериальной оптимизации методом свётртки критериев.
+Подобные серии задач возникают, если задача глобальной оптимизации имеет дискретный параметр или при решении задачи многокритериальной оптимизации методом свёртки критериев.
 Рассматриваемый алгоритм использует отображения типа кривой Пеано для редукции многомерных задач оптимизации к одномерным.
@@
 
-\keywords{Computational mathematics / Вычислительная математика, Parallel and distributed computing technologies / Технологии параллельных и распределенных вычислений}
+\keywords{Глобальная оптимизация, параллельные вычисления, алгоритмы прямой оптимизации, равномерная сходимость}
 
@@
 зачастую оказывается сложнее, чем локальная оптимизация в тысячемерном пространстве. Для последней может оказаться достаточно
-применения простейшего метода градиентного спуска, в то время как чтобы \textit{гаранитрованно} отыскать глобальный оптимум методам
+применения простейшего метода градиентного спуска, в то время как чтобы \textit{гарантированно} отыскать глобальный оптимум методам
 оптимизации приходится накапливать информацию о поведении целевой функции во всей области поиска \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}.
@@
 экстремума было оценено примерно с одинаковым качеством. Обычно серию из \(q\) задач решают либо последовательно, либо
-параллельно порциями по \(p,\:p<<q\) задач, где \(p\) --- количество параллельных вычислительных усройств.
+параллельно порциями по \(p,\:p<<q\) задач, где \(p\) --- количество параллельных вычислительных устройств.
 Такой подход ведёт к тому, что в каждый момент времени до окончания вычислений
@@
 механизмом, порождающим наборы задач заданной размерности с заданным количеством нелинейных ограничений \cite{GergelBarkalov2019}.
-Кроме искуственно сгенерированных задач, рассматриваемый метод протестирован также
+Кроме искусственно сгенерированных задач, рассматриваемый метод протестирован также
 на множестве задач, возникающем при решении задачи многокритериальной оптимизации
@@
 \section{Постановка задачи глобальной оптимизации}
-В рамках данной работы будем рассмативать следующую постановку задачи глобальной
+В рамках данной работы будем рассматривать следующую постановку задачи глобальной
 оптимизации: найти глобальный минимум \(N\)-мерной функции \(\varphi(y)\) в гиперинтервале
@@
 в процессе которых были проведены испытания в \(k\) точках \(x_i, 1\leqslant i\leqslant k\).
-Тогда точкa \(x^{k+1}\) поисковых испытаний следующей \((k+1)\)-ой
+Тогда точка \(x^{k+1}\) поисковых испытаний следующей \((k+1)\)-ой
 итерации определяются в соответствии с правилами:
@@
 
-Параллельная модификация метода не отличается от рассматриаемой в \cite{BarkalovStrongin2018}
+Параллельная модификация метода не отличается от рассматриваемой в \cite{BarkalovStrongin2018}
 и заключается в выборе \(p\) интервалов на шаге 6 и выполнения \(p\) испытаний параллельно
@@
 
-Вместе с системой GCGen распростаняются примеры её использования и построения
+Вместе с системой GCGen распространяются примеры её использования и построения
 наборов задач, каждая из которых состоит из целевой функции и двух ограничений,
@@
 В сочетании с GCGen было порождено 100 задач размерности 2. Каждая из задач имеет два ограничения.
-Также с целью демонстарции того, что свойства метода сохраняются при существенно разных свойствах задач
+Также с целью демонстрации того, что свойства метода сохраняются при существенно разных свойствах задач
 был сгенерирован смешанный класс, состоящий из 50 задач с функциями GKLS и 50 задач с функциями \(F_{GR}\).
-На рис. \ref{fig:isolines} предствалены примеры линий уровня рассматриваемых задач. Допустимая область отмечена зелёным цветом.
+На рис. \ref{fig:isolines} представлены примеры линий уровня рассматриваемых задач. Допустимая область закрашена.
 
@@
 При оценке качества метода и его реализации кроме ускорения от распараллеливания и времени выполнения также будем принимать во внимание среднее максимальное расстояния (в смысле \(l_{\inf}\)-нормы) текущей оценки оптимума до его реального положения,
-вычисленное на множестве задач (\ref{eq:many_problems}): \(D_{avg}\) и \(D_{max}\). Динамика этих величи в процессе оптимизации
+вычисленное на множестве задач (\ref{eq:many_problems}): \(D_{avg}\) и \(D_{max}\). Динамика этих величин в процессе оптимизации
 показывает, насколько равномерно метод распределяет ресурсы между задачами.
@@
 Из таблицы видно, что ускорение по итерациям \(S_i\) растёт линейно с увеличением числа потоков \(p\),
-в то время, как укорение по времени \(S_p\) увеличивается не так быстро, что говорит о неидеальной
+в то время, как ускорение по времени \(S_p\) увеличивается не так быстро, что говорит о неидеальной
 реализации алгоритма. Увеличить реальное ускорение, верхней границей для которого является
@@
 кривая Парето имеет вогнутости, что не соответствует решению, указанному в \cite{BinhKorn1999} и
-означает нехватку ресурсов для решении вспомогательных задач.
+означает нехватку ресурсов для решения некоторых из вспомогательных задач.
 Для оценки качества решения также был вычислен показатель \(Spacing(SP)\) \cite{RiquelmeLucken2015},
@@
 показатели параллельного ускорения по времени. Также планируется реализовать версию
-рассматриваемого алгоритма, работающего на распределённой памяти по схеме, описанной в \cite{BarkalovLebedev2017}.
+рассматриваемого алгоритма, работающего на распределённой памяти по схеме, описанной в \cite{BarkalovLebedev2017_2}.
 

### 2019-12-02 2020/PAVT_Sovrasov/pavt_paper.tex
PAVT: Add results on 3d problems

@@
 Генератор GKLS \cite{Gaviano2003} позволяет получать функции заданной размерности и с заданным количеством экстремумов.
-В сочетании с GCGen было порождено 100 задач размерности 2. Каждая из задач имеет два ограничения.
+В сочетании с GCGen были порождены два множества по 100 задач размерности 2 и 3. Каждая из задач имеет два ограничения.
 Также с целью демонстрации того, что свойства метода сохраняются при существенно разных свойствах задач
-был сгенерирован смешанный класс, состоящий из 50 задач с функциями GKLS и 50 задач с функциями \(F_{GR}\).
+был сгенерирован смешанный класс, состоящий из 50 задач с двухмерными функциями GKLS и 50 задач с функциями \(F_{GR}\).
 На рис. \ref{fig:isolines} представлены примеры линий уровня рассматриваемых задач. Допустимая область закрашена.
@@
 Результаты решения тестовых задач последовательной и параллельной версией модифицированного ИАГП
-для решения множества задач представлены в таблице \ref{tab:speedup}. Для всех классов задач параметр \(r=4.7\).
+для решения множества задач представлены в таблице \ref{tab:speedup}. Для всех двухмерных классов задач параметр \(r=4.7\).
+В случае трёхмерных задач \(r=4.7,\: \varepsilon_\nu=0.1\).
 Во всех экспериментах в целевые функции и ограничения была внесена дополнительная вычислительная нагрузка так,
@@
     \hline
-    GKLS-based \
+    GKLS \& \(F_{GR}\) based \
+      & 1 & 51434 & 90.20 & -    & - \\
+      & 2 & 25698 & 56.96 & 2.00 & 1.58 \\
+      & 4 & 13015 & 36.67 & 3.95 & 2.46 \\
+      & 6 & 8332  & 26.85 & 6.17 & 3.36 \\
+    \hline
+    GKLS based 2d \
       & 1 & 59066 & 97.53 & -    & - \\
@@
     \hline
-    GKLS \& \(F_{GR}\) based \
-         & 1 & 51434 & 90.20 & -    & - \\
-         & 2 & 25698 & 56.96 & 2.00 & 1.58 \\
-         & 4 & 13015 & 36.67 & 3.95 & 2.46 \\
-         & 6 & 8332  & 26.85 & 6.17 & 3.36 \\
-        \noalign{\smallskip}\hline
+    GKLS based 3d \
+      & 1 & 782544 & 1117.55 & -    & - \\
+      & 2 & 397565 & 752.92  & 1.97 & 1.48 \\
+      & 4 & 208073 & 526.67  & 3.76 & 2.12 \\
+      & 6 & 142089 & 445.45  & 5.50 & 2.51 \\
+    \hline
   \end{tabular}
@@
 Не смотря на значительную разницу в структуре задач, модифицированный ИАГП
-значительно быстрее уменьшает как максимальное, так и среднее отклонение оценок от оптимумов.
+гораздо быстрее уменьшает как максимальное, так и среднее отклонение оценок от оптимумов.
 Это говорит о наличии равномерной сходимости по всему множеству совместно решаемых задач.

### 2019-12-02 2020/PAVT_Nested_Full/pavt_paper.tex
Include two more references

@@
 the global optimization method), and at the lower one the local optimization problems are solved (using 
-the local method). The generated local subproblems will not affect each other, and the solving of these 
+the local method). The generated local subproblems will not affect each other, and solving of these 
 ones can be performed in parallel. The numerical experiments have been conducted on several 
@@
 the minimum interaction energy are relevant \cite{Posypkin2014}. In particular,  such problems arise 
-in the development of new medical drugs \cite{Sulimov}. The global optimization problems arise in 
-many other applications as well (see the review \cite{Pinter2006}).
+in the development of new medical drugs \cite{Sulimov}. 
+Another example is the problem of finding optimal control parameters for complex processes \cite{Kalyulin2017,Modorskii2016}.
+The global optimization problems arise in many other applications as well (see the review \cite{Pinter2006}).
 

### 2019-12-10 2019/Automation/Barkalov.tex
Comments for reviewer

@@
 \documentclass[12pt]{a&t}
-%\usepackage{hyperref}
+\usepackage{hyperref}
 \usepackage{graphicx}
@@
 \begin{equation}\label{block_nested_i}
-\varphi^i(u_1,...,u_i) = \min_{u^{i+1}\in D_{i+1}} \varphi_{i+1}(u_1,...,u_i,u
+\varphi_i(u_1,...,u_i) = \min_{u^{i+1}\in D_{i+1}} \varphi_{i+1}(u_1,...,u_i,u
 _{i+1}), 1\leq i\leq M-1.
@@
 Настоящая работа предлагает обобщение адаптивной схемы для случая многомерных подзадач.
-Дадим краткое описание ее основных элементов.
+
+Перед изложением новой схемы организации вычислений еще раз отметим, что в рамках исходной многошаговой схемы (или же ее блочного варианта) порождаемые подзадачи решаются строго последовательно; получаемая в результате иерархическая схема порождения и решения подзадач (в виде дерева) представлена на рис.
+\ref{fig0}.
+Построение этого дерева происходит динамически при решении исходной задачи (\ref{main_problem}).
+Вычисление значения функции $\varphi_i(y_1,y_2,...,y_i)$  на $i$-м уровне требует полного решения всех задач одного из поддеревьев уровня $i+1$.
+При этом листьями дерева задач являются значения целевой функции $\varphi(y)=\varphi(y_1,y_2,...,y_N)$, вычисляемые в процессе решения задачи (\ref{main_problem}).
+
+Адаптивная многошаговая схема редукции размерности предполагает устранение последовательного решения задач в соответствии с их иерархией в дереве задач. В рамках адаптивной схемы:
+\begin{itemize}
+	\item 
+для вычисления значения функции $i$-го уровня из (\ref{block_nested_i}) порождается новая задача уровня $i+1$, в которой проводится только одно испытание, после чего новая порожденная задача включается в множество уже имеющихся задач, подлежащих решению;
+	\item 
+	итерация глобального поиска состоит в выборе одной задачи из множества имеющихся задач, в которой проводится одно испытание; точка проведения нового испытания определяется в соответствии с алгоритмом из раздела \ref{SectionPeano}.
+	\item
+в качестве минимальных значений функций из (\ref{block_nested_i}) используются их текущие оценки, полученные на основе накопленной поисковой информации.
+\end{itemize}
+
+Краткое описание основных шагов блочной адаптивной схемы редукции размерности приведено ниже.
 
@@
 Завершение процесса оптимизации происходит, когда в корневой задаче выполняется условие остановки алгоритма, решающего эту задачу.
+
+Новые элементы предлагаемой схемы редукции размерности состоят в следующем.
+\begin{itemize}
+	\item 
+В адаптивной многошаговой схеме все порождаемые задачи решаются совместно -- на каждой итерации выбирается одна (наиболее перспективная) подзадача из множества подзадач, в которой проводится одно испытание.
+Такой подход, с одной стороны, требует одновременного хранения поисковой информации во всех решаемых подзадачах.
+С другой стороны, он позволяет (при необходимости) увеличивать необходимую точность решения задачи (\ref{main_problem}) без проведения повторного решения всех подзадач (\ref{block_nested_i}).
+Например, начальная стадия оптимизации может быть выполнена с достаточно грубой точностью, что позволит быстро получить некоторую оценку решения; далее (после анализа результатов) точность может быть повышена и процесс глобального поиска может быть продолжен.
+	\item
+	В исходной многошаговой схемой редукции размерности решение каждой возникающей подзадачи проводится ``до конца'', т.е.
+порождения новой поздадачи не произойдет, пока не будет решена текущая подзадача (пусть даже и не содержащая решения исходной многомерной задачи).
+В адаптивной схеме подобные бесперспективные подзадачи могут быть отбракованы уже на начальном этапе их решения.
+\item
+Наличие правила выбора подзадачи при выполнения очередной итерации в рамках адаптивной схемы позволяет получать различные ее варианты, соответствующие разным правилам выбора подзадачи.
+Например, как исходная многошаговая схема редукции размерности, так и схема редукции размерности с помощью единственной кривой Пеано могут быть получены как частные случаи нового подхода.
+\item
+Существование множества одновременно решаемых подзадач позволяет организовать их параллельное решение и эффективно задействовать ресурсы современных вычислительных систем.
+Некоторые результаты в данном направлении представлены в \cite{Gergel19}.
+\end{itemize}
 
@@
 
+\bibitem{Gergel19}
+{\it Gergel V., Grishagin V., Israfilov R.} Parallel dimensionality reduction for multiextremal optimization problems // LNCS. 2019. V. 11657. P. 166--178.
+
 \bibitem{Gaviano03}

### 2019-12-11 2019/Automation/Barkalov.tex
Fix reviewrs question

@@
 \begin{equation}\label{block_nested_i}
-\varphi_i(u_1,...,u_i) = \min_{u^{i+1}\in D_{i+1}} \varphi_{i+1}(u_1,...,u_i,u
+\varphi_i(u_1,...,u_i) = \min_{u_{i+1}\in D_{i+1}} \varphi_{i+1}(u_1,...,u_i,u
 _{i+1}), 1\leq i\leq M-1.
@@
 
-Перед изложением новой схемы организации вычислений еще раз отметим, что в рамках исходной многошаговой схемы (или же ее блочного варианта) порождаемые подзадачи решаются строго последовательно; получаемая в результате иерархическая схема порождения и решения подзадач (в виде дерева) представлена на рис.
-\ref{fig0}.
-Построение этого дерева происходит динамически при решении исходной задачи (\ref{main_problem}).
-Вычисление значения функции $\varphi_i(y_1,y_2,...,y_i)$  на $i$-м уровне требует полного решения всех задач одного из поддеревьев уровня $i+1$.
-При этом листьями дерева задач являются значения целевой функции $\varphi(y)=\varphi(y_1,y_2,...,y_N)$, вычисляемые в процессе решения задачи (\ref{main_problem}).
+Перед изложением новой схемы организации вычислений еще раз отметим, что в рамках исходной многошаговой схемы (или же ее блочного варианта) порождаемые подзадачи решаются строго последовательно; получаемая в результате иерархическая схема порождения и решения подзадач (в виде дерева) представлена на рис.~\ref{fig0}.
+Построение этого дерева происходит динамически в процессе решения исходной задачи (\ref{main_problem}).
+При этом вычисление одного значения функции $\varphi_i(y_1,y_2,...,y_i)$ на $i$-м уровне требует полного решения всех задач одного из поддеревьев уровня $i+1$.
 
-Адаптивная многошаговая схема редукции размерности предполагает устранение последовательного решения задач в соответствии с их иерархией в дереве задач. В рамках адаптивной схемы:
+Адаптивная многошаговая схема редукции размерности изменяет порядок решения подзадач: они будут решаться не по одной (в соответствии с их иерархией в дереве задач), а одновременно, т.е.
+будет существовать некоторое множество подзадач, находящихся в процессе решения.
+В рамках новой схемы:
 \begin{itemize}
@@
 	\item 
-	итерация глобального поиска состоит в выборе одной задачи из множества имеющихся задач, в которой проводится одно испытание; точка проведения нового испытания определяется в соответствии с алгоритмом из раздела \ref{SectionPeano}.
+	итерация глобального поиска состоит в выборе одной (наиболее перспективной) задачи из множества имеющихся задач, в которой проводится одно испытание; точка проведения нового испытания определяется в соответствии с алгоритмом из раздела \ref{SectionPeano};
 	\item
@@
 
-Краткое описание основных шагов блочной адаптивной схемы редукции размерности приведено ниже.
-
+Краткое описание основных шагов блочной адаптивной схемы редукции размерности состоит в следующем.
 Пусть вложенные подзадачи вида (\ref{block_nested_i}) решаются с помощью алгоритма глобального поиска, описанного в разделе \ref{SectionPeano}.
@@
 Поэтому на каждой итерации выбирается подзадача с максимальной характеристикой для проведения в ней очередного испытания.
-Это испытание либо вычисляет значение целевой функции $\varphi(y)$ (если выбранная подзадача принадлежала уровню $j=M$), либо порождает новые подзадачи согласно (\ref{block_nested_i}) при $j\leq M-1$.
+Это испытание либо приводит к вычислению значения целевой функции $\varphi(y)$ (если выбранная подзадача принадлежала уровню $j=M$), либо порождает новые подзадачи согласно (\ref{block_nested_i}) при $j\leq M-1$.
 В последнем случае новые порожденные задачи добавляются к текущему множеству задач, вычисляются их характеристики и процесс повторяется.
@@
 
-Новые элементы предлагаемой схемы редукции размерности состоят в следующем.
+Отметим новые элементы предлагаемой схемы редукции размерности.
 \begin{itemize}
@@
 	\item
-	В исходной многошаговой схемой редукции размерности решение каждой возникающей подзадачи проводится ``до конца'', т.е.
-порождения новой поздадачи не произойдет, пока не будет решена текущая подзадача (пусть даже и не содержащая решения исходной многомерной задачи).
+	В исходной многошаговой схемой редукции размерности решение каждой возникающей подзадачи проводится <<до конца>>, т.е.
+порождения новой поздадачи не произойдет, пока не будет решена текущая подзадача.
+Но точка $\overline{u_i}$ ее решения, вообще говоря, не будет совпадать с $i$-й компонентой вектора решения $y^* = (u_1^\ast,...,u_i^\ast,...,u_M^\ast)$ исходной многомерной задачи.
 В адаптивной схеме подобные бесперспективные подзадачи могут быть отбракованы уже на начальном этапе их решения.
@@
 
-\begin{figure}[ht]
+\begin{figure}
 \begin{minipage}{0.5\linewidth}

### 2020-01-03 2019/NUMTA_Barkalov/Special Issure/article.tex
Added formulas, literature.

@@
 \usepackage{lineno,hyperref}
+\usepackage[T1]{fontenc}
 \modulolinenumbers[5]
@@
 
-\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
-\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}
+\title{Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints}
 
 %% Group authors per affiliation:
-\author{Elsevier\fnref{myfootnote}}
-\address{Radarweg 29, Amsterdam}
-\fntext[myfootnote]{Since 1880.}
-
-%% or include affiliations in footnotes:
-\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
-\ead[url]{www.elsevier.com}
-
-\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
-\ead{support@elsevier.com}
-
-\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
-\address[mysecondaryaddress]{360 Park Avenue South, New York}
+\author{Roman Strongin}
+\author{Konstantin Barkalov}
+\author{Semen Bevzuk}
+%\address[mymainaddress]{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia}
+%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
+%\cortext[mycorrespondingauthor]{Corresponding author}
+%\ead{support@elsevier.com}
+\address{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia}
 
 \begin{abstract}
-This template helps you to create a properly formatted \LaTeX\ manuscript.
+In the present work, the constrained global optimization problems, in which the functions are of the “black box” type and satisfy the Lipschitz condition, are considered.
+The algorithms for solving the problems of this class require the use of the adequate estimates of the a priori unknown Lipschitz constants for the problem functions.
+A novel approach presented in this paper is based on a simultaneous use of two estimates of the Lipschitz constant: an overestimated and the underestimated ones.
+The upper estimate provides the global convergence whereas the lower one reduces the number of trials necessary to find the global optimizer with required precision.
+At that, the considered algorithm of solving the constrained problems doesn’t use the ideas of the penalty function method; each constraint of the problem is accounted for separately.
+The convergence conditions of the proposed algorithm are formulated according to corresponding theorem.
+The results of the numerical experiments on the series of the multiextremal problems with the non-convex constraints demonstrating the efficiency of the proposed scheme of dual Lipschitz constant estimates are presented.
 \end{abstract}
@@
 \begin{keyword}
-\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
-\MSC[2010] 00-01\sep  99-00
+Global optimization, Multiextremal problems, Non-convex constraints, Lipschitz constant estimates 
 \end{keyword}
@@
 
-\section{The Elsevier article class}
+\section{Introduction}
 
-\paragraph{Installation} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.
+\section{Index Method of Accounting for the Constraints and Dimensionality Reduction}
 
-\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript.
-Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal.
-It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.
+\begin{equation}\label{problem}
+	\min{\left\{\varphi(y):y \in D, \; g_j(y)\leq 0, \; 1 \leq j \leq m\right\}},
+\end{equation}
 
-\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class.
-In addition, it features commands and options to format the
-\begin{itemize}
-\item document style
-\item baselineskip
-\item front matter
-\item keywords and MSC codes
-\item theorems, definitions and proofs
-\item lables of enumerations
-\item citation style and labeling.
-\end{itemize}
+\begin{equation}\label{D}
+	D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.
+\end{equation}
 
-\section{Front matter}
+\begin{equation}\label{lipschitz_condition}
+	\left|g_j(y')-g_j (y'')\right| \leq L_j \|y'-y''\|, \; y', y''\in D, \; 1\leq j \leq m+1.
+\end{equation}
 
-The author names and affiliations could be formatted in two ways:
-\begin{enumerate}[(1)]
-\item Group the authors per affiliation.
-\item Use footnotes to indicate the affiliations.
-\end{enumerate}
-See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.
+\begin{equation}\label{epsilon_reserved_solution}
+	\varphi(y_{\epsilon})=\min{\left\{\varphi(y):y \in D, \; g_j(y)\leq -{\epsilon}_j, \; 1 \leq j \leq m\right\}},
+\end{equation}
 
-\section{Bibliography styles}
+\begin{equation}\label{Y_epsilon}
+	Y_{\epsilon}=\left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
+\end{equation}
+
+\begin{equation}\label{D_sets}
+	D_1 = D, \; D_{j+1} = \left\{ y \in D_j: \; g_j(y) \leq 0 \right\}, \; 1 \leq j \leq m.
+\end{equation}
+
+\begin{equation}\label{one_dimensional_problem}
+	\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\},
+\end{equation}
+
+\[
+	\left|g_j(y(x'))-g_j(y(x''))\right| \leq K_j \left|x'-x'' \right|^{1/N}, \; x', x''\in [0,1], \; 1\leq j \leq m+1,
+\]
+
+\begin{equation}\label{Q_intervals}
+	Q_1=[0,1], \; Q_{j+1}=\left\{x \in Q_j : g_j(y(x)) \leq 0 \right\}, \; 1 \leq j \leq m.
+\end{equation}
+
+\[
+	g_j(y(x)) \leq 0, \; 1 \leq j \leq \nu-1, \; g_{\nu}(y(x))>0,
+\]
+
+\begin{equation}\label{trial_result}
+	x^k, \; z^k = g_{\nu}\left( y(x^k) \right), \; \nu = \nu(x^k),
+\end{equation}
+
+\begin{equation}\label{reduction_problem}
+	\psi(x^*)=\min \left\{\psi(x): x \in [0,1] \right\},
+\end{equation}
+
+$$\[
+	\psi(x)=\frac{g_{\nu}(y(x))}{K_{\nu}} - 
+	\left\{
+   \begin{array}{lr}
+     0, & \nu \leq m,\\
+     \frac{\varphi^*}{K_{m+1}}, & \nu = m + 1.
+   \end{array}
+	\right
+\]$$
+
+\[
+	\psi(x)=\frac{\varphi(y(x))-\varphi^*}{K_{m+1}}
+\]
+
+\section{Index Algorithm of Global Search}
+
+\[
+	0=x_0<x_1<\dots <x_k<x_{k+1}=1,
+\]
+
+\[
+	I_\nu =\left\{i:1 \leq i \leq k, \; \nu=\nu(x_i) \right\}, \; 1 \leq \nu \leq m+1,
+\]
+
+\[
+	M=\max\left\{\nu=\nu(x_i), \; 1 \leq i \leq k \right \}.
+\]
+
+\begin{equation}\label{current_lower_bounds}
+	\mu_{\nu} = \max\left\{ \frac{\left|z_i-z_j\right|}{ (x_i - x_j)^{1/N} }, \; i,j \in I_\nu, \; i>j \right\},
+\end{equation}
+
+\begin{equation}\label{z_estimates}
+	z_\nu^\ast = \left\{
+   \begin{array}{lr}
+     -\epsilon_\nu, & \nu < M,\\
+     \min\{ z_i: i\in I_\nu \}, & \nu = M,
+   \end{array}
+	\right.
+\end{equation}
+
+\begin{equation}\label{R_1}
+	R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
+\end{equation}
+\begin{equation}\label{R_2}
+	R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
+\end{equation}
+\begin{equation}\label{R_3}
+R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
+\end{equation}
+
+\begin{equation}\label{MaxR}
+	R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
+\end{equation}
+
+\[
+	x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
+\]
+
+\[
+	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
+\]
+
+\[
+	g_j \left( y(x) \right) = G_j \left( y(x) \right), \; x \in Q_j, \; 1 \leq j \leq m+1,
+\]
+
+\begin{equation}\label{theorem_inequalities}
+	r_{\nu}\mu_{\nu} > 2^{3-\frac{1}{N}}L_{\nu}\sqrt{N+3}, \; 1 \leq \nu \leq m+1, 
+\end{equation}
+
+\[
+	\varphi(\bar y) = \inf\left\{\varphi(y^q):g_j(y^q) \leq 0, 1 \leq j \leq m, q = 1,2,\ldots \right\} \leq \varphi(y_{\epsilon}).
+\]
+
+\begin{equation}\label{epsilon_nu}
+	\epsilon_{\nu} = \mu_{\nu}\delta, \; 1 \leq \nu \leq m, 
+\end{equation}
+
+\section{Index Algorithm with Dual Lipschitz Constant Estimates}
+
+\begin{equation}\label{estimates_Lipschitz_constants}
+	\frac{r_{\nu}\mu_{\nu}}{2^{3-\frac{1}{N}}L_{\nu}\sqrt{N+3}}, \; 1 \leq \nu \leq m+1, 
+\end{equation}
+
+\begin{equation}\label{R_glob_1}
+	R(i)=R_{glob}(i)=R_{loc}(i)=2\Delta_i>0. 
+\end{equation}
+
+\[
+	z_i+z_{i-1}-2z_M^* = |z_i-z_{i-1}|\leq \mu_M\Delta_i
+\]
+
+\begin{eqnarray}
+	R(i) = \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{z_i+z_{i-1}-2z_M^*}{r_M\mu_M} \geq \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\mu_M\Delta_i}{r_M\mu_M} \nonumber \\
+	\geq \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\Delta_i}{r_M} = \Delta_i + \Delta_i\frac{(z_i-z_{i-1})^2/\Delta_i^2}{r_M^2\mu_M^2\Delta_i^2} - 2\frac{\Delta_i}{r_M}
+\nonumber \\
+	\geq \Delta_i + \frac{\Delta_i}{r_M^2} - 2\frac{\Delta_i}{r_M} = \Delta_i \left( 1-\frac{1}{r_M} \right)^2>0
+	\nonumber
+\end{eqnarray}
+
+\begin{equation}\label{R_glob_greater_R_loc}
+	R_{glob}(i)=\Delta_i\left(1-\frac{1}{r_M^{glob}} \right)^2 > \Delta_i\left(1-\frac{1}{r_M^{loc}} \right)^2 = R_{loc}(i)
+\end{equation}
+
+\begin{equation}\label{R_glob_equal_R_loc}
+	R_{glob}(i) = \rho R_{loc}(i)
+\end{equation}
+
+\begin{equation}\label{R_max}
+	R(i) = \max{\left\{ R_{glob}(i), \rho R_{loc}(i) \right\}}
+\end{equation}
+
+\begin{equation}\label{R_max_t}
+	R(t) = \max{\left\{ R(i): 1 \leq i \leq k+1\right\}}
+\end{equation}
+
+\[
+	x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
+\]
+
+\[
+	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
+\]
+
+\begin{eqnarray}
+	\varphi(y_1, y_2)=-1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}- \nonumber \\
+	-\left(0.5(y_1-1)(y_2-1)\right)^4\exp{\left\{2-\left(0.5(y_1-1)\right)^4-(y_2-1)^4\right\}}
+	\nonumber
+\end{eqnarray}
+
+\begin{eqnarray}
+	g_1(y_1, y_2)= 0.01 \left( (y_1-2.2)^2+(y_2-1.2)^2-2.25 \right) \leq 0, \nonumber \\
+	g_2(y_1, y_2)= 100 \left(1-(y_1-2)^2/1.44+(0.5y_2)^2 \right) \leq 0, \nonumber \\
+	g_3(y_1, y_2)= 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0
+	\nonumber
+\end{eqnarray}
+
+\begin{equation}\label{phi_bar_y}
+	\varphi(\bar y) = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \leq \varphi(y_{\epsilon}).
+\end{equation}
+
+\[
+	R_{glob}(t) < \rho R_{loc}(t)
+\]
+
+\[
+	z_j=g_{\nu}\left( y(x_j) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{1/N}, \; \nu=\nu(x_j)
+\]
+
+\[
+	z_{j-1}=g_{\nu}\left( y(x_{j-1}) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \; \nu=\nu(x_{j-1})
+\]
+
+\[
+	g_{\nu}\left( y(\bar x) \right) \leq -\epsilon_{\nu}, \; 1\leq\nu\leq m.
+\]
+
+\[
+	g_{m+1}\left( y(\bar x) \right) \leq z_{m+1}^*
+\]
+
+\begin{eqnarray}
+	R(j) = \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^*}{r_{\nu}\mu_{\nu}} \geq \nonumber \\
+	\Delta_i + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+	=\Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+	=\Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+	\geq \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+	= \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \geq 0
+	\nonumber
+\end{eqnarray}
+
+\begin{eqnarray}
+	R(j) = 2\Delta_i - 4\frac{z_j-z_{\nu}^*}{r_{\nu}\mu_{\nu}} \geq 2\Delta_i+4\frac{z_{\nu}^*-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+	= 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+	\geq 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0
+	\nonumber
+\end{eqnarray}
+
+\begin{equation}
+	R_{glob}(j) > \rho R_{loc}(t)
+\end{equation}
+
+\section{TEST}
 
@@
 
-Here are two sample references: \cite{Feynman1963118,Dirac1953888}.
+Here are two sample references:
+
+\cite{Evtushenko, Piyavskii1972, Strongin1970, Shubert1972, Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Pinter1996, Jones2009, Wood1991, Meewella1988, Mladineo1986, Vaz2009, Stripinis2019, Paulavicius2016, Pillo2012, Pillo2016, Barkalov2017_1, Barkalov2017_2, Sergeyev2006, Zilinskas2008, Sovrasov2019}.
 

### 2020-01-09 2019/NUMTA_Barkalov/Special Issure/article.tex
Some minor changes to compile LaTeX

@@
 
-$$\[
+\begin{eqnarray}
 	\psi(x)=\frac{g_{\nu}(y(x))}{K_{\nu}} - 
@@
    \end{array}
-	\right
-\]$$
+	\right.
+\end{eqnarray}
 
@@
 
-\cite{Evtushenko, Piyavskii1972, Strongin1970, Shubert1972, Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Pinter1996, Jones2009, Wood1991, Meewella1988, Mladineo1986, Vaz2009, Stripinis2019, Paulavicius2016, Pillo2012, Pillo2016, Barkalov2017_1, Barkalov2017_2, Sergeyev2006, Zilinskas2008, Sovrasov2019}.
+\cite{Evtushenko, Piyavskii1972, Shubert1972, Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Pinter1996, Jones2009, Wood1991, Meewella1988, Mladineo1986, Vaz2009, Stripinis2019, Paulavicius2016, Pillo2012, Pillo2016, Barkalov2017_1, Barkalov2017_2, Sergeyev2006, Zilinskas2008, Sovrasov2019}.
 

### 2020-01-09 2019/NUMTA_Barkalov/Special Issure/article.tex
Added left alignment in formulas.
Merged the bibliography

@@
 \begin{eqnarray}
-	R(i) = \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{z_i+z_{i-1}-2z_M^*}{r_M\mu_M} \geq \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\mu_M\Delta_i}{r_M\mu_M} \nonumber \\
-	\geq \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\Delta_i}{r_M} = \Delta_i + \Delta_i\frac{(z_i-z_{i-1})^2/\Delta_i^2}{r_M^2\mu_M^2\Delta_i^2} - 2\frac{\Delta_i}{r_M}
+	R(i) & = & \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{z_i+z_{i-1}-2z_M^*}{r_M\mu_M} \geq \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\mu_M\Delta_i}{r_M\mu_M} \nonumber \\
+	& \geq & \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\Delta_i}{r_M} = \Delta_i + \Delta_i\frac{(z_i-z_{i-1})^2/\Delta_i^2}{r_M^2\mu_M^2\Delta_i^2} - 2\frac{\Delta_i}{r_M}
 \nonumber \\
-	\geq \Delta_i + \frac{\Delta_i}{r_M^2} - 2\frac{\Delta_i}{r_M} = \Delta_i \left( 1-\frac{1}{r_M} \right)^2>0
+	& \geq & \Delta_i + \frac{\Delta_i}{r_M^2} - 2\frac{\Delta_i}{r_M} = \Delta_i \left( 1-\frac{1}{r_M} \right)^2>0  
 	\nonumber
@@
 \begin{eqnarray}
-	g_1(y_1, y_2)= 0.01 \left( (y_1-2.2)^2+(y_2-1.2)^2-2.25 \right) \leq 0, \nonumber \\
-	g_2(y_1, y_2)= 100 \left(1-(y_1-2)^2/1.44+(0.5y_2)^2 \right) \leq 0, \nonumber \\
-	g_3(y_1, y_2)= 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0
+	g_1(y_1, y_2) &=& 0.01 \left( (y_1-2.2)^2+(y_2-1.2)^2-2.25 \right) \leq 0, \nonumber \\
+	g_2(y_1, y_2) &=& 100 \left(1-(y_1-2)^2/1.44+(0.5y_2)^2 \right) \leq 0, \nonumber \\
+	g_3(y_1, y_2) &=& 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0
 	\nonumber
@@
 \begin{eqnarray}
-	R(j) = \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^*}{r_{\nu}\mu_{\nu}} \geq \nonumber \\
-	\Delta_i + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	=\Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	=\Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	\geq \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	= \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \geq 0
+	R(j) &=& \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^*}{r_{\nu}\mu_{\nu}}  \nonumber \\
+	&\geq& \Delta_i + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+	&=& \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+	&=& \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+	&\geq& \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+	&=& \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \geq 0
 	\nonumber
@@
 \begin{eqnarray}
-	R(j) = 2\Delta_i - 4\frac{z_j-z_{\nu}^*}{r_{\nu}\mu_{\nu}} \geq 2\Delta_i+4\frac{z_{\nu}^*-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	= 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	\geq 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0
+	R(j) &=& 2\Delta_i - 4\frac{z_j-z_{\nu}^*}{r_{\nu}\mu_{\nu}} \geq 2\Delta_i+4\frac{z_{\nu}^*-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+	&=& 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+	&\geq& 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0
 	\nonumber
@@
 
-\cite{Evtushenko, Piyavskii1972, Shubert1972, Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Pinter1996, Jones2009, Wood1991, Meewella1988, Mladineo1986, Vaz2009, Stripinis2019, Paulavicius2016, Pillo2012, Pillo2016, Barkalov2017_1, Barkalov2017_2, Sergeyev2006, Zilinskas2008, Sovrasov2019}.
-
-\section*{References}
+\cite{Evtushenko1971, Piyavskii1972, Shubert1972, Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Pinter1996, Jones2009, Wood1991, Meewella1988, Mladineo1986, Vaz2009, Stripinis2019, Paulavicius2016, Pillo2012, Pillo2016, Barkalov2017_1, Barkalov2017_2, Sergeyev2006, Zilinskas2008, Sovrasov2019}.
 

### 2020-01-10 2019/NUMTA_Barkalov/Special Issure/article.tex
Add some references

@@
 \usepackage{lineno,hyperref}
-\usepackage[T1]{fontenc}
+%\usepackage[T1]{fontenc}
 \modulolinenumbers[5]
@@
 
-\cite{Evtushenko1971, Piyavskii1972, Shubert1972, Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Pinter1996, Jones2009, Wood1991, Meewella1988, Mladineo1986, Vaz2009, Stripinis2019, Paulavicius2016, Pillo2012, Pillo2016, Barkalov2017_1, Barkalov2017_2, Sergeyev2006, Zilinskas2008, Sovrasov2019}.
+\cite{Evtushenko1971, Piyavskii1972, Shubert1972, Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Pinter1996, Jones2009, Wood1991, Meewella1988, Mladineo1986, Vaz2009, Stripinis2019, Paulavicius2016, Pillo2012, Pillo2016, Barkalov2017_1, Barkalov2017_2, Sergeyev2006, Zilinskas2008, Sovrasov2019, Kvasov2003, Sergeyev2010,Sergeyev2016, Horst1996, Gablonsky2001, Jones1993, Gaviano2003, Barkalov2018, Paulavicius2014, Sergeyev2015, Strongin2018,Gergel2017_2,Gergel2019}.
 

### 2020-01-10 2019/NUMTA_Barkalov/Special Issure/article.tex
Added Strongin1970

--- 2020-01-09 2019/NUMTA_Barkalov/Special Issure/article.tex
@@
 
-\cite{Evtushenko1971, Piyavskii1972, Shubert1972, Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Pinter1996, Jones2009, Wood1991, Meewella1988, Mladineo1986, Vaz2009, Stripinis2019, Paulavicius2016, Pillo2012, Pillo2016, Barkalov2017_1, Barkalov2017_2, Sergeyev2006, Zilinskas2008, Sovrasov2019}.
+\cite{Evtushenko1971, Piyavskii1972, Shubert1972, Strongin1970, Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Pinter1996, Jones2009, Wood1991, Meewella1988, Mladineo1986, Vaz2009, Stripinis2019, Paulavicius2016, Pillo2012, Pillo2016, Barkalov2017_1, Barkalov2017_2, Sergeyev2006, Zilinskas2008, Sovrasov2019, Kvasov2003, Sergeyev2010,Sergeyev2016, Horst1996, Gablonsky2001, Jones1993, Gaviano2003, Barkalov2018, Paulavicius2014, Sergeyev2015, Strongin2018, Gergel2017_2, Gergel2019}.
 

### 2020-01-14 2020/PAVT_Sovrasov/pavt_paper.tex
PAVT: fix missing comma
PAVT: fix comments from reviewers

@@
 \authors{В.В.~Соврасов}
-\organizations{Нижегородский государственный университет им. Н.И. Лобачевского}
+\organizations{Нижегородский государственный университет им. Н.И.~Лобачевского}
 
@@
 Решение серии таких задач при ограниченных вычислительных
-ресурсах является ещё более сложной проблемой: помимо поиска глобального экстремума необходимо
+ресурсах является ещё более сложной задачей: помимо поиска глобального экстремума необходимо
 распределять вычислительные ресурсы так, чтобы сразу во всех решаемых задачах положение глобального
@@
 реализации алгоритма. Увеличить реальное ускорение, верхней границей для которого является
-\(S_i\) возможно путём оптимизации взаимодействий между копиями ИАГП и это планируется сделать в ходе будущей работы.
+\(S_i\), возможно путём оптимизации взаимодействий между копиями ИАГП и это планируется сделать в ходе будущей работы.
 
@@
 
-В ходе работы была реализована поддержка нелинейных ограничений в алгоритме, решающим
+В ходе работы была реализована поддержка нелинейных ограничений в алгоритме, решающeм
 множество задач глобальной оптимизации. Проведены численные эксперименты, демонстрирующие

### 2020-01-15 2019/NUMTA_Barkalov/Special Issure/article.tex
Add Introduction

@@
 
-\journal{Journal of \LaTeX\ Templates}
+%\journal{Journal of \LaTeX\ Templates}
 
@@
 \begin{abstract}
-In the present work, the constrained global optimization problems, in which the functions are of the “black box” type and satisfy the Lipschitz condition, are considered.
-The algorithms for solving the problems of this class require the use of the adequate estimates of the a priori unknown Lipschitz constants for the problem functions.
-A novel approach presented in this paper is based on a simultaneous use of two estimates of the Lipschitz constant: an overestimated and the underestimated ones.
-The upper estimate provides the global convergence whereas the lower one reduces the number of trials necessary to find the global optimizer with required precision.
-At that, the considered algorithm of solving the constrained problems doesn’t use the ideas of the penalty function method; each constraint of the problem is accounted for separately.
-The convergence conditions of the proposed algorithm are formulated according to corresponding theorem.
-The results of the numerical experiments on the series of the multiextremal problems with the non-convex constraints demonstrating the efficiency of the proposed scheme of dual Lipschitz constant estimates are presented.
+This paper considers the constrained global optimization problems, in which the functions are of the “black box” type and satisfy the Lipschitz condition.
+The algorithms for solving the problems of this class require the use of adequate estimates of the a priori unknown Lipschitz constants for the problem functions.
+A novel approach presented in this paper is based on a simultaneous use of two estimates of the Lipschitz constant: an overestimated and an underestimated one.
+The upper estimate provides the global convergence whereas the lower one reduces the number of trials necessary to find the global optimizer with the required accuracy.
+In this case,  the considered algorithm for solving the constrained problems doesn’t use the ideas of the penalty function method; each constraint of the problem is accounted for separately.
+The convergence conditions of the proposed algorithm are formulated in the corresponding theorem.
+The results of the numerical experiments on a series of multiextremal problems with non-convex constraints demonstrating the efficiency of the proposed scheme of dual Lipschitz constant estimates are presented.
 \end{abstract}
@@
 \section{Introduction}
+	In this paper, we consider constrained global optimization problems and the numerical methods for solving problems of this type.
+Such problems are often encountered in applications (see, for example, the review \cite{Pinter2006}).
+As a rule, problem functions are defined by a program code, i.e.
+they are “black-box” functions, for which the computing of the values can be a computationally costly operation since in applied problems it will require  numerical modeling.
+Any  possibility to estimate reliably the global optimum in the multiextremal problems with “black-box” functions is based fundamentally on the auxiliary assumptions relating the possible values of the problem functions to their known values at the points of the trials already performed.
+
+	The assumption that the problem functions satisfy the Lipschitz condition is typical for many applied problems (in this case, one can speak about the Lipschitzian optimization problems) because the ratio of the functions’ variations to corresponding ones of the variables usually cannot exceed some threshold determined by the limited energy of variations in the system.
+This threshold can be described using the Lipschitz constant.
+
+	The methods for solving the Lipschitzian optimization problems constitute a very important area in the development of the global optimization methods and are the objects of investigations of many researchers.
+The first algorithms of this class were proposed as early as in the 1970s \cite{Evtushenko1971, Piyavskii1972, Shubert1972, Strongin1970}; since that time, this area has been continuing to develop extensively (see, for example \cite{Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Jones2009}).
+Note that these global optimization algorithms are, as a rule, oriented towards solving either  unconstrained optimization problems
+	\cite{Zilinskas, Sergeyev, Pinter, Jones}
+or problems with constraints of a simple structure \cite{Vaz2009, Paulavicius2016}.
+The solving of the problems with complex non-convex constraints is usually performed with the use of the penalty functions method \cite{Stripinis2019, Pillo2012, Pillo2016}, which has a number of drawbacks.
+In particular, this method cannot be applied to the problems with the partially defined functions (i.e.
+when any constraint is violated, the values of all the other functions of the problem remain undefined).
+
+	This situation is typical of many optimal design problems, when, in case of violation of some conditions of the modeled object's functioning, other characteristics of the object appear to be undefined.
+For example, the modeled object is an electronic device, and its response time is minimized.
+However, the response time is undefined if at a given combination of parameters the device fails to operate.
+Such a partial computability of the functions in the constrained optimization problems complicates essentially (in some cases, makes impossible) the application of the penalty functions method.
+
+	Within the framework of our research, an original approach to the minimization of the multiextremal functions with non-convex constraints called the index method of the accounting for the constraints
+	\cite{Strongin} 
+was used.
+This approach is based on separate accounting for each constraint of the problem and does not involve the use of the penalty functions.
+According to the rules of the index method, every iteration called trial at the corresponding point of the search domain includes a sequential checking of the feasibility of the problem constraints at this point.
+When the first violation of any constraint is found, the trial is terminated (the values of the remaining problem functions are not computed at this point) and the transition to the next iteration point is initiated.
+This allows solving the problems with partially defined functions.
+The experimental comparison of the index method of accounting for the constraints and the penalty function method demonstrated the advantages of the index method in terms of the number of iterations and the number of computations of the problem functions \cite{Barkalov2017_1, Barkalov2017_2}.
+
+	Note that the values of the Lipschitz constant for the problem functions are usually not known a priori, which  makes their estimation one of the key problems in the construction of the Lipschitzian optimization methods.
+The methods utilizing the values of the Lipschitz constant predefined a priori (see, for example, \cite{Piyavskii1972, Shubert1972, Wood1991, Meewella1988, Mladineo1986}) are important in the theoretical aspect, but it is difficult to apply such methods in applied problems (in which the information of the Lipschitz constant values is absent).
+
+		The value of the unknown Lipschitz constant estimate affects the convergence rate of the algorithm essentially.
+Therefore, the issue of its correct estimate is very important.
+An underestimate of the true value of the constant may lead to the loss of the algorithm convergence to the global solution.
+At the same time, too high a value of the constant estimate not corresponding to real behavior of the function results in slow convergence of the algorithm to the global minimum point.
+
+		Several typical methods for adaptive estimation of the Lipschitz constant according to the results of the performed search trial are known:
+\begin{itemize}[\labelitemii]
+  \item global estimate of the constant $L$ in the whole search domain $D$ \cite{Horst1996, Pinter1996, Strongin2000}.
+  \item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ \cite{Kvasov2003, Sergeyev2010, Sergeyev2016}.
+	\item choice of the estimates of the constant $L$ from some set of possible values \cite{Gablonsky2001, Jones1993, Jones2009, Sergeyev2006}.
+\end{itemize}
+
+	Each of the approaches listed above has its own advantages and drawbacks.
+For example, the use of the global estimate only in the entire search domain may slow down the convergence of the algorithm to the global minimum point.
+On the other hand, the use of the local estimates accelerating the convergence requires an adequate adjustment of the algorithm parameters in order to ensure the global convergence.
+
+	In this paper, we consider an algorithm, in which it is proposed to use two global estimates of the Lipschitz constant, one of them being much greater than the other one.
+The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
+The choice of one of the two estimates to be used in the rules of the algorithm is performed in the adaptive manner, depending on the behavior of the function.
+This work continues the research reported in the paper
+	\cite{NUMTA2019},
+in which the preliminary results have been obtained for the unconstrained problems.
+
+	The main part of the paper has the following structure.
+In Section 2, the index method of accounting for the constraints and the dimensionality reduction scheme for the constrained global optimization problems are described.
+In Section 3, a description of the computational rules of the index global search algorithm is given and its theoretical properties are formulated.
+In Section 4, the index algorithm with dual Lipschitz constant estimates is formulated and its convergence is proved.
+Section 5 presents the results of the numerical experiments carried out on several series of the multiextremal problems with non-convex constraints.
+Section 6 concludes the paper.
 

### 2020-01-15 2019/NUMTA_Barkalov/Special Issure/article.tex
Add Index Method of Accounting for the Constraints and Dimensionality Reduction

@@
 \section{Index Method of Accounting for the Constraints and Dimensionality Reduction}
-
+	We consider the global optimization problems of the form
 \begin{equation}\label{problem}
@@
 \end{equation}
-
 \begin{equation}\label{D}
@@
 
+	It is supposed that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_{j}(y), 1 \leq j \leq m$, of the constraints satisfy the Lipschitz condition
 \begin{equation}\label{lipschitz_condition}
@@
 \end{equation}
+with \textit{a priori} unknown constant $L_j, 1\leq j \leq m+1$, and may be multiextremal.
 
+	The analytical form of the functions $g_{j}(y)$ may be unknown, i.e.
+these functions may be defined algorithmically (“black-box” functions).
+In this case, it is supposed that even a single search trial may be a computationally costly operation since in applied problems it involves the need to perform numerical modeling.
+
+	Along with the exact solution $y^\ast$ of the problem (\ref{problem}), we will consider also the $\varepsilon$-reserved solution $y_{\epsilon}$, for which 
 \begin{equation}\label{epsilon_reserved_solution}
@@
 \end{equation}
-
+and the set of solutions $Y_{\epsilon}$, which are not worse than the $\varepsilon$-reserved solution in the objective function value, i.e.
 \begin{equation}\label{Y_epsilon}
@@
 \end{equation}
+Fig. \ref{fig:eps_reserved_solution} illustrates the concepts introduced.
 
+\begin{figure}[h]
+	\center{\includegraphics[width=0.7\linewidth]{img/epsilon_reserved_solution}}
+	\caption{$\varepsilon$-reserved solution}
+	\label{fig:eps_reserved_solution}
+\end{figure}
+
+	The existence of the $\varepsilon$-reserved solutions of problem (\ref{problem}) can be interpreted as some analog of the regularity condition in classical nonlinear programming problems.
+The applied role of this condition should be also noted.
+Even if the exact solution $y^\ast$ is found, its practical implementation is possible as some approximation of $y^\ast$ only.
+Therefore, it is important that there are feasible points from $Q_{m+1}$ close enough to $y^\ast$ (in coordinates and values).
+The existence of the $\varepsilon$-reserved solution guarantees the existence of such points.
+In this case, the domain $Y_{\epsilon}$ may play the role of a set of suitable approximations.
+
+	We will consider the initial problem (\ref{problem}) assuming that the problem functions $g_{j}(y), 1 \leq j \leq m+1$ may be defined only partly.
+It means that they are defined and calculable only in the sets
 \begin{equation}\label{D_sets}
@@
 \end{equation}
+This situation is typical for the optimal design problems when some conditions for the functioning of the modeled object are not met and its other characteristics appear to be undefined.
+In this approach, the first violation of any constraint indicating that the point ybeing considered is no longer feasible terminates further computations at the point $y$.
 
+	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions \cite{Sergeyev2006} or simplicial partitions \cite{Zilinskas2008}).
+In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve).
+Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
 \begin{equation}\label{one_dimensional_problem}
@@
 \end{equation}
+The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000, Sergeyev2013}.
+Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$ where the number $m>1$ is a parameter of the construction (the density of the evolvent).
+In this case, the construction density of the evolvent is determined by the accuracy of the possible practical implementation of the problem solution found.
+For  illustration, Fig.
+\ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] are shown for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
+\begin{figure}[h]
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=0.8\linewidth]{img/evolvent_N=2_m=4} \\ a)}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=0.8\linewidth]{img/evolvent_N=2_m=5} \\ b)}
+	\end{minipage}
+	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
+	\label{fig:evolvent}
+\end{figure}
 
-\[
+	The dimensionality reduction scheme considered here matches a multidimensional problem with Lipschitzian minimized function and Lipschitzian constraints to a one-dimensional problem, in which the corresponding functions satisfy the uniform H\"{o}lder condition (see \cite{Strongin2000, Sergeyev2013}), i.e.
+$$
 	\left|g_j(y(x'))-g_j(y(x''))\right| \leq K_j \left|x'-x'' \right|^{1/N}, \; x', x''\in [0,1], \; 1\leq j \leq m+1,
-\]
+$$
+where $N$ is the dimensionality of the initial multidimensional problem and the coefficients $K_j$ are related to the Lipschitz constants $L_j$ of the initial problem by the relations $K_j \leq 2L_j \sqrt{(N+3}$.
 
+	According to (\ref{D_sets}) and (\ref{one_dimensional_problem}), the reduced one-dimensional functions $g_j (y(x)), 1 \leq j \leq m+1$ will also be defined and calculable in the corresponding intervals
 \begin{equation}\label{Q_intervals}
@@
 \end{equation}
+These relations allow introducing a classification of the points $x \in [0,1]$ according to the number $\nu = \nu(x)$ of the constraints calculable at the given point.
+We will call this number $\nu = \nu(x)$ the \textit{index} of the point $x$.
+The index can be determined from the conditions
+$$
+	g_j(y(x)) \leq 0, \; 1 \leq j \leq \nu-1, \; g_{\nu}(y(x))>0,
+$$
+where the last inequality is insufficient if $\nu = m+1$.
 
-\[
-	g_j(y(x)) \leq 0, \; 1 \leq j \leq \nu-1, \; g_{\nu}(y(x))>0,
-\]
+	Thus, a search trial at some point $x^k \in [0,1]$ performed at the $k^{th}$ iteration of the algorithm will consist of the following sequence of operations:
+\begin{enumerate} 
+  \item Determine the image $y^k=y(x^k )$ with the use of the evolvent $y(x)$. 
+  \item Compute the values $g_1 (y^k ), \ldots ,g_{\nu} (y^k )$ where $\nu = \nu(x)$ is the index of the point.
+	\item Fix the result of the trial as three values 
+\end{enumerate}
 
@@
 
+	According to this scheme, the trial is terminated at the point $y^k=y(x^k)$ when the first violated constraint  is found.
+
+	In the case when the point $y^k$ is a feasible one, i.e.
+when $y(x^k) \in Q_(m+1)$, the trial involves checking all constraints and computing the objective function.
+Here, the value of the index is accepted to be equal to the quantity $\nu = m+1$.
+
+	The main idea of the index scheme of accounting for the constraints consists in the reduction of the constrained problem (\ref{one_dimensional_problem}) (the functions of which satisfy the H\"{o}lder condition with corresponding coefficients $K_j$) to an unconstrained problem of the form
 \begin{equation}\label{reduction_problem}
@@
 \end{equation}
-
 \begin{eqnarray}
@@
      0, & \nu \leq m,\\
-     \frac{\varphi^*}{K_{m+1}}, & \nu = m + 1.
+     \frac{\varphi^\ast}{K_{m+1}}, & \nu = m + 1.
    \end{array}
@@
 \end{eqnarray}
+where $\nu = \nu(x)$ and $\varphi^\ast$ is the objective function value at the point of the problem solution (\ref{one_dimensional_problem}).
 
-\[
-	\psi(x)=\frac{\varphi(y(x))-\varphi^*}{K_{m+1}}
-\]
+	The value $\psi(x)$ at $\nu(x) \leq m$ matches the value of the left-hand part of the first constraint violated at this point $g_{\nu}(y(x))$ divided by the constant $K_{\nu}$.
+In the case when $\nu(x) = m+1$
+$$
+	\psi(x)=\frac{\varphi(y(x))-\varphi^\ast}{K_{m+1}}
+$$
+and, consequently, the set of the points, in which the function $\psi(x)$ turns to zero matches the set of the global minimum points of the problem (\ref{one_dimensional_problem}).
+It should be noted that in any subdomain $Q_j$ the function $\psi(x)$ has the H\"{o}lder coefficient equal to 1.
 
-\section{Index Algorithm of Global Search}
+	These considerations constitute the basis of the index global search algorithm, according to which the unknown constants $K_{\nu}$ and the unknown minimum value $\varphi^\ast$ are substituted in the course of computations by their current approximations.
+
+\section{Index-Based Global Search Algorithm}
 

### 2020-01-15 2019/NUMTA_Barkalov/Special Issure/article.tex
Add Index-Based Global Search Algorithm

@@
 \end{equation}
-\begin{eqnarray}
+\begin{eqnarray} \nonumber
 	\psi(x)=\frac{g_{\nu}(y(x))}{K_{\nu}} - 
@@
 \section{Index-Based Global Search Algorithm}
+	The index-based global search algorithm under consideration for solving the problem (\ref{problem}) (here it is given according to \cite{Strongin2000}) involves the construction of a sequence of points $x^k \in (0,1)$, at which the search trials are conducted with the results of the type (\ref{trial_result}).
+The set of triplets $\{x^k, z^k=g_{\nu}\big(y(x^k)\big), \nu=\nu(x^k)\}, 1 \leq k \leq n$  makes up the search information accumulated by the method after performing $n$ steps.
+The operation of the index-based global search algorithm is defined by the following rules.
 
-\[
+	The first trial is executed at an arbitrary internal point $x^1 \in (0,1)$. The selection of the point $ x^{k+1}, k \geq 1$, of any next trial is determined by the following steps. \\
+Step 1. Renumber the points $x^1, \ldots, x^k$ of the preceding trials by the lower indices in the increasing order of the coordinate values, i.e.
+$$
 	0=x_0<x_1<\dots <x_k<x_{k+1}=1,
-\]
+$$
+and juxtapose to them the values $z_i=g_{\nu}\big(y(x_i)\big),\nu=\nu(x_i), 1 \leq i \leq k$, computed at these points.
+The points $x_0=0$ and $x_{k+1}=1$ are introduced additionally and the values $z_0$ and $z_{k+1}$ are not defined.
+\\
+Step 2. Classify the indices $i, 1 \leq i \leq k$, of the trial points in accordance with the number of the constraints of the problem fulfilled at these points by constructing the sets
+$$
+	I_\nu =\left\{i:1 \leq i \leq k, \; \nu=\nu(x_i) \right\}, \; 1 \leq \nu \leq m+1,
+$$
+including the numbers of all points $x_i, 1 \leq i \leq k$ with the same values of $\nu$.
+The end points $x_0=0$ and $x_{k+1}=1$ are interpreted as the ones having the indices equal to zero.
+An additional set $I_0=\{0, k+1\}$ corresponds to these points.
 
-\[
-	I_\nu =\left\{i:1 \leq i \leq k, \; \nu=\nu(x_i) \right\}, \; 1 \leq \nu \leq m+1,
-\]
-
-\[
+	Determine the maximum value of the index 
+$$
 	M=\max\left\{\nu=\nu(x_i), \; 1 \leq i \leq k \right \}.
-\]
-
+$$ \\
+Step 3. Compute the current lower bounds
 \begin{equation}\label{current_lower_bounds}
@@
 \end{equation}
-
+for the unknown H\"{o}lder constants $K_{\nu}$ of the functions $g_{\nu}, 1 \leq \nu \leq m+1$.
+If the set $I_{\nu}$ includes less than two elements or if $\mu_{\nu}$ appears to be equal to zero, then set $\mu_{\nu}=1$.
+\\
+Step 4. For all nonempty sets $I_{\nu}, 1 \leq \nu \leq m+1$, compute the estimates
 \begin{equation}\label{z_estimates}
@@
 \end{equation}
-
+where the nonnegative numbers $(\epsilon_1,\ldots,\epsilon_m )$ are the algorithm parameters. \\
+Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$, compute the \textit{characteristics} $R(i)$:
 \begin{equation}\label{R_1}
@@
 \end{equation}
-
+where $\Delta_i=(x_i-x_{i-1})^{1/N}$.
+The values $r_{\nu} > 1, 1 \leq \nu \leq m+1$, are the parameters of the algorithm.
+With an appropriate selection of $r_{\nu}$ one can use the product $r_{\nu}\mu_{\nu}$ as an estimate of H\"{o}lder constants $K_{\nu}, 1 \leq \nu \leq m+1$.
+\\
+Step 6. Find the interval $(x_{t-1}, x_t)$ with the maximum characteristic
 \begin{equation}\label{MaxR}
 	R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
-\end{equation}
+\end{equation} \\
+Step 7. Execute the next trial at the middle point of the interval $(x_{t-1}, x_t)$ if the indices of the points $x_{t-1}, x_t$ are not the same, i.e.,
+$$
+	x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
+$$
+Otherwise, execute a trial at the point
+$$
+	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
+$$
+The inequality $\Delta_t \leq \epsilon$ can be taken as the termination condition, where $t$ is from (\ref{MaxR}) and $\epsilon>0$ is the predefined accuracy.
 
-\[
-	x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
-\]
-
-\[
-	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
-\]
-
-\[
+	The convergence conditions for the index global search algorithm can be formulated in the form of the following theorem from \cite{Strongin2000}. 
+\\
+\textbf{Theorem}. Assume that the following conditions are satisfied:
+\begin{enumerate} 
+  \item Problem (\ref{problem}) has an $\varepsilon$-reserved solution $y_{\varepsilon}$ from (\ref{epsilon_reserved_solution}).
+  \item Functions $g_j(y), 1 \leq j \leq m+1$ admit Lipschitzian (with the constants $L_j$) extensions $G_{j}(y)$ over the whole feasible domain $D$ from (\ref{D}), i.e.
+$$
 	g_j \left( y(x) \right) = G_j \left( y(x) \right), \; x \in Q_j, \; 1 \leq j \leq m+1,
-\]
-
+$$
+where the sets $Q_j$ are from (\ref{Q_intervals}), and $y(x)$ is the curve from (\ref{one_dimensional_problem}).
+	\item The parameters of the method $\epsilon_{\nu}, 1 \leq \nu \leq m$ used in the rule (\ref{z_estimates}) are the corresponding components of the reserve vector $\epsilon_{R}$ from (\ref{epsilon_reserved_solution}).
+	\item For a sufficiently large iteration number $q$, the values $\mu_{\nu}$ from (\ref{current_lower_bounds}) satisfy the inequalities
 		\begin{equation}\label{theorem_inequalities}
@@
 		\end{equation}
-
-\[
+Then, any limit point $\bar{y}$ of the sequence of the trials $\{y^k\}=\big\{y\big(x^k\big)\big\}$ generated by the index algorithm for problem (\ref{problem}) belongs to the set $Y_{\epsilon}$ form (\ref{Y_epsilon}) and satisfies the conditions
+$$
 	\varphi(\bar y) = \inf\left\{\varphi(y^q):g_j(y^q) \leq 0, 1 \leq j \leq m, q = 1,2,\ldots \right\} \leq \varphi(y_{\epsilon}).
-\]
-
+$$
+\end{enumerate}
+\textbf{Note 1}.
+The computational scheme of the algorithm is described in the form convenient for  formulating its theoretical properties.
+The software implementation of this scheme may be made more frugal.
+For example, according to Step 7, the point $x^{k+1}$ of the scheduled trial will be an internal point of the interval $(x_{t-1}, x_t)$, i.e.
+$x^{k+1} \in (x_{t-1}, x_t)$.
+Therefore, the set of trial points $x^1,\ldots, x^k$ (and the related trial results (\ref{trial_result})) may not be ordered  at every step of the algorithm but may be maintained in an ordered state using appropriate data structures.
+\\
+\textbf{Note 2}.
+The use of the set of parameters $\epsilon_1,\ldots,\epsilon_m$ (the values of which correspond to the coordinates of the reserve vector $\epsilon_R$) in the index algorithm will significantly reduce the density of the trial sequence in the vicinity of the feasible domain boundaries.
+However, this raises a question about the particular values of the reserve vector components $\epsilon_R$.
+In \cite{Strongin2000}, it was proposed to choose the components of the vector $\epsilon_R$ according to the formula
 \begin{equation}\label{epsilon_nu}
@@
 \end{equation}
+where $\mu_{\nu}$ are the current estimates of the H\"{o}lder constant from (\ref{current_lower_bounds}) and $\delta>0$ is a parameter.
+This approach to choosing the reserve vector components allows one to take into account the properties of different problem functions characterized by their Lipschitz constants.
+This method for choosing the reserve vector was used in the numerical experiments, the results of which are presented below.
+\\
+\textbf{Note 3}.
+The index algorithm for solving the constrained problems is based on the information-statistical algorithm of global search (AGS) for solving unconstrained problems.
+AGS has been compared repeatedly with both deterministic and nature-inspired global optimization algorithms.
+The results of the experimental comparison (see, for example, \cite{Sovrasov2019}) show that AGS is superior to many known methods of similar purpose, including DIRECT and DIRECT\textit{l}.
+
+	At the same time, the index scheme of accounting for the constraints also demonstrated its advantages (in the number of computed values of the problem functions) over the traditional penalty functions method in solving a wide range of model problems (see the results of the experimental comparison in \cite{Barkalov2017_1, Barkalov2017_2}.)
 

### 2020-01-16 2019/NUMTA_Barkalov/Special Issure/article.tex
Add Index Algorithm with Dual Lipschitz Constant Estimates

@@
 $$
-where $N$ is the dimensionality of the initial multidimensional problem and the coefficients $K_j$ are related to the Lipschitz constants $L_j$ of the initial problem by the relations $K_j \leq 2L_j \sqrt{(N+3}$.
+where $N$ is the dimensionality of the initial multidimensional problem and the coefficients $K_j$ are related to the Lipschitz constants $L_j$ of the initial problem by the relations
+\begin{equation}\label{K_leq_L}
+	K_j \leq 2L_j \sqrt{N+3}, \; 1\leq j \leq m+1.
+\end{equation}
 
@@
 \section{Index Algorithm with Dual Lipschitz Constant Estimates}
+	The index global search algorithm described in the previous section is intended for solving  the multiextremal problems, in which the objective function and the functions of constraints satisfy the Lipschitz condition.
+In this case, \textit{a priori} estimates of the constant values are not required for the algorithm to work.
+These estimates are computed in the course of global search, based on the available search information.
 
+	According to Theorem 1 from Section 3, the sequence of trial points $\{y^k\}$ will converge to a global minimizer $y^\ast$ if the condition (\ref{theorem_inequalities}) is satisfied.
+So far, an appropriate choice of the parameters $r_{\nu}$ from (\ref{R_2}) allows using the values
 \begin{equation}\label{estimates_Lipschitz_constants}
@@
 \end{equation}
+as the estimates of the Lipschitz constants for the problem functions $g_{j}(y), 1 \leq j \leq m+1$.
 
+	Compliance with condition (\ref{estimates_Lipschitz_constants}) will be achieved when choosing a high value of the parameters $r_{\nu}$.
+However, in this case the constant will be overestimated, and the method will perform a large number of trials until the stopping condition is fulfilled.
+
+	The choice of low values of the parameters $r_{\nu}$ (which corresponds to an underestimate of the Lipschitz constant) would reduce the number of trials considerably but could destroy the convergence to the global extremum (due to possible violation of the condition (\ref{theorem_inequalities})).
+
+	An approach, in which two estimates of the Lipschitz constant for each function are used in the rules of the algorithm, is promising.
+One of these estimates is significantly greater than the other, which corresponds to the use of two sets of parameters $r_{\nu}^{glob}$ and $r_{\nu}^{loc}$ in the algorithm where $r_{\nu}^{glob}>r_{\nu}^{loc}>1, 1 \leq \nu \leq m+1$.
+When using the parameter $r_{\nu}^{loc}$ we shall deal with the smaller estimate of the Lipschitz constant, and using the parameter $r_{\nu}^{glob}$ will correspond to the larger one.
+
+	We will refer to the characteristics of the $i^{th}$ interval as the global ones $(R_{glob}(i))$ if the parameters $r_{\nu}=r_{\nu}^{glob}$ were used when computing the characteristic according to (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) and as the local ones  $(R_{loc}(i))$ if the parameters $r_{\nu}=r_{\nu}^{loc}$ were used.
+
+	The key idea of the novel method is computing two characteristics for each interval and balancing (based on these two) the local and global search.
+The local search provides a new trial in the vicinity of the current point $y_k^\ast$  whereas during the global search, a new trial could be performed in a subdomain located far away from $y_k^\ast$.
+
+	Note that Theorem 1 on the convergence of the index method implies the generation of an infinite sequence of trials $\{y^k\}=\big\{y(x^k )\big\}$ by the algorithm.
+In this case, at the final stage of the search (i.e.
+in the case when one of the points of this trial sequence falls into the vicinity of the $\varepsilon$-reserved solution $y^\ast$) there always exists an interval, at one of the boundaries of which the current minimum value of the objective function $z_{m+1}^\ast$  from (\ref{z_estimates}) is achieved.
+It is this interval that will have the highest characteristic since it is the most promising one for performing further trials.
+
+	Let us establish the following important property for the global and local characteristics of an interval containing the current optimal point.
+
+	\textbf{Statement}.
+Let us assume that the current minimum value $z_M^\ast$  from (\ref{z_estimates}) at the $k^{th}$ iteration is achieved at the boundary point of some interval $(x_{i-1}, x_i), i=i(k)$.
+Then, $R_{glob}(i) \geq R_{loc}(i)$.
+
+	\textbf{Proof}. Let us consider two possible cases of the values of the indices $\nu(x_{i-1})$ and $\nu(x_i)$. 
+\begin{enumerate} 
+  \item If $\nu(x_{i-1}) \neq \nu(x_i)$, then according to  formulae (\ref{R_2}), (\ref{R_3}) 
 \begin{equation}\label{R_glob_1}
@@
 \end{equation}
-
-\[
+  \item If $\nu(x_{i-1}) = \nu(x_i) = M$, then, taking into account the rules (\ref{current_lower_bounds}), the following inequality will hold
+$$
 	z_i+z_{i-1}-2z_M^* = |z_i-z_{i-1}|\leq \mu_M\Delta_i
-\]
-
+$$
+\end{enumerate}
+From here, it will follow that for any value $r_M>1$ the following inequalities will hold
 \begin{eqnarray}
@@
 \end{eqnarray}
-
+And because $r_{M}^{glob} > r_{M}^{loc} > 1$ then
 \begin{equation}\label{R_glob_greater_R_loc}
@@
 \end{equation}
+The statement is proved.
 
+	This statement can be interpreted as the incomparability of the local and global characteristics: the global characteristic of the best interval is always larger.
+At the same time, it follows from the formulae (\ref{R_glob_1}), (\ref{R_glob_greater_R_loc}) that
 \begin{equation}\label{R_glob_equal_R_loc}
-	R_{glob}(i) = \rho R_{loc}(i)
+	R_{glob}(i) = \rho R_{loc}(i),
 \end{equation}
+where $\rho=\Big(\frac{1-1/r_{M}^{glob}}{1-1/r_{M}^{loc}}\Big)^2$, if $\nu(x_{i-1})=\nu(x_i)=M$, and $\rho=1$, if $\nu(x_{i-1}) \neq \nu(x_i)$.
+In the case of the multiplication by the coefficient $\rho$, the local and global characteristics of the search intervals become comparable, and one can choose the best (i.e.
+the largest) one.
 
+	Based on the relation (\ref{R_glob_equal_R_loc}), one can formulate an \textit{index algorithm with dual estimate of the Lipschitz constants} (hereafter referred to as IA-DL).
+
+	The first four steps of the algorithm repeat Step 1 -- Step 4 of the initial index algorithm from Section 3.
+	Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$ 
+\begin{flushleft}
+	\begin{itemize}[\labelitemii]
+		\item Compute the global characteristics $R_{glob}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{glob}$
+		\item Compute the local characteristics $R_{loc}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{loc}$
+		\item Determine the characteristic $R(i)$ as
 \begin{equation}\label{R_max}
@@
 \end{equation}
+where $\rho=\Big(\frac{1-1/r_{M}^{glob}}{1-1/r_{M}^{loc}}\Big)^2$, if $\nu(x_{i-1})=\nu(x_i)=\nu$, and $\rho=1$, if $\nu(x_{i-1}) \neq \nu(x_i)$.
+	\end{itemize}
+\end{flushleft}
 
+	Step 6. Find the interval $(x_{t-1}, x_t)$ with the maximum characteristic
 \begin{equation}\label{R_max_t}
@@
 \end{equation}
+and fix the value $r_{\nu}=r_{\nu}^{loc}$ if $\rho R_{loc}(t) > R_{glob}(t)$, otherwise fix $r_{\nu}=r_{\nu}^{glob}$.
 
-\[
+	Step 7. Execute the next trial at the middle point of the interval $(x_{t-1}, x_t)$ if the indices of the points $x_{t-1}$, $x_t$ are not the same, i.e.
+$$
 	x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
-\]
+$$
+Otherwise, execute a trial at the poin
+$$
+	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
+$$
+When computing the point $x^{k+1}$, use the value of the parameter $r_{\nu}$ fixed at Step 6.
+The algorithm terminates if the condition $\Delta_{t} \leq \epsilon$ is satisfied where $t$ is from (\ref{R_max_t}) and $\epsilon>0$ is the predefined accuracy.
 
-\[
-	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
-\]
-
+	Let us illustrate the work of the algorithms IA and IA-DL when minimizing the function
 \begin{eqnarray}
@@
 \end{eqnarray}
-
+in the range $0 \leq y_1 \leq 4$, $1 \leq y_2 \leq 3$ subject to the constraints
 \begin{eqnarray}
@@
 \end{eqnarray}
+The feasible domain of this problem is not connected and consists of three non-convex subdomains.
+The global minimum $\varphi(y^\ast)=-1.489$ is achieved at the point $y^\ast =(0.942, 0.944)$ (the values are given with the accuracy of $10^{-3}$).
 
+\begin{figure}[h]
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=\linewidth]{img/stronginC3_IA} \\ a)}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=\linewidth]{img/stronginC3_IA-DL} \\ b)}
+	\end{minipage}
+	\caption{Solving a problem by IA (a) and IA-DL (b) methods}
+	\label{fig:example_strongin}
+\end{figure}
+
+	In Fig.
+\ref{fig:example_strongin}, the search domain (square) is shown, the feasible domain is highlighted by green, and the trial points generated by IA (a) and IA-DL (b) are marked.
+The best point is marked by red.
+For both methods, the evolvent with the parameter $m=10$ was used, the accuracy of the search for the solution $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ from (\ref{epsilon_nu}).
+For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$ were used.
+The same values were used for the IA-DL method as the parameters $r_{\nu}^{glob}, 1 \leq \nu \leq 4$.
+Also, the parameters $r_{\nu}^{loc}=1.5, 1 \leq \nu \leq 4$ were used.
+To solve the problem with a predefined accuracy, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method.
+The results of larger-scale numerical experiments are presented in Section 4.
+
+	Let us formulate and prove the convergence conditions for the new algorithm IA-DL.\\
+\textbf{Theorem 2}.
+Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1$ the conditions of Theorem 1 are fulfilled.
+Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1, 1\leq \nu \leq m+1$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e.
+satisfies the condition
 \begin{equation}\label{phi_bar_y}
@@
 \end{equation}
+\textbf{Proof}.
 
-\[
+	Because for the IA with the parameters $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1$ the conditions of Theorem 1 are fulfilled, then any limit point $\bar{y}=y(\bar{x})$ of the trial sequence generated by the IA for the problem (\ref{problem}) satisfies the conditions (\ref{phi_bar_y}).
+
+	Let us show that $\bar y$ is also the limit point of the sequence of trials $\{y^k\}=\{y(x^k)\}$, generated by the IA-Dl, and the sequence $\{y^k\}$ has no new limit points.
+Let us carry out the proof by contradiction.
+Assume that $\bar y$ is not a limit point of the sequence $\{y^k\}$ and there exists a limit point $y' \neq \bar y$.
+
+	At every iteration of the IA-DL method there will exist such interval $[x_{j-1}, x_j], j=j(k)$ that $\bar x \in [x_{j-1}, x_j]$.
+According to our assumption that $\bar x$ is not the limit point, there will exist such index $k$, starting from which the trial points will not fall into the interval $[x_{j-1}, x_j]$.
+That is, to conduct a new trial at the point $x^{k+1} \in [x_{t-1}, x_t]$ at every iteration such interval $[x_{t-1}, x_t], t=t(k)$ will be selected that $t \neq j$.
+
+	Therefore, in the formula (\ref{R_max}) for computing the characteristic, the second component takes the highest values, i.e. 
+$$
 	R_{glob}(t) < \rho R_{loc}(t)
-\]
+$$
+Then, there will exist a limit point $x' \neq \bar x$, at that either $\nu(x') \leq m$ (i.e.
+the point $y(x')$ is not a feasible one) or $z_{m+1}^\ast =g_{m+1}(y(x'))>g_{m+1}(y(\bar x))$ (i.e.
+the value of the objective function at the limit point $y(x')$ is greater than at the point $y(\bar x)$).
 
-\[
+	Let us show that if the conditions of Theorem 1 are satisfied, the inequality $R_{glob}(j)>0$ will hold. According to the second condition of Theorem 1 and to the condition (\ref{K_leq_L}), 
+$$
 	z_j=g_{\nu}\left( y(x_j) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{1/N}, \; \nu=\nu(x_j)
-\]
+$$
+$$
+	z_{j-1}=g_{\nu}\left( y(x_{j-1}) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \; \nu=\nu(x_{j-1})
+$$
 
-\[
-	z_{j-1}=g_{\nu}\left( y(x_{j-1}) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \; \nu=\nu(x_{j-1})
-\]
+	Since the limit point $\bar u = y(\bar x)$ belongs to the set $Y_{\epsilon}$, then
+$$
+	g_{\nu}\left( y(\bar x) \right) \leq -\epsilon_{\nu}, \; 1\leq\nu\leq m.
+$$
+$$
+	g_{m+1}\left( y(\bar x) \right) \leq z_{m+1}^*
+$$
 
-\[
-	g_{\nu}\left( y(\bar x) \right) \leq -\epsilon_{\nu}, \; 1\leq\nu\leq m.
-\]
-
-\[
-	g_{m+1}\left( y(\bar x) \right) \leq z_{m+1}^*
-\]
-
+	Then, the characteristic $R_{glob}(j)$ of the interval $[x_{j-1}, x_j], j=j(k)$ will be strictly positive, i.e.
+$R_{glob}(j)>0$.
+Indeed, at $\nu=\nu(x_j)=\nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
 \begin{eqnarray}
@@
 \end{eqnarray}
-
+At $\nu=\nu(x_j ) > \nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
 \begin{eqnarray}
@@
 
+	The case of $\nu=\nu(x_{j-1})>\nu(x_j)$ is treated similarly and leads to the same inequality $R(j)>0$.
+
+	Since we have assumed that $x'$ is a limit point for the IM-DL and  $x' \neq \bar x$, then there exists such a sequence of the intervals $[x_{t-1}, x_t], t=t(k)$ that $x' \in [x_{t-1}, x_t], R_{loc}(t)>0$, and $R_{loc}(t) \rightarrow 0$.
+This property of the characteristics of the intervals shrinking to the limit point has been established when proving the theorem on the convergence of the IA in \cite{Sergeyev2013}.
+
+	Therefore, for large enough, the following inequality will be true
 \begin{equation}
@@
 \end{equation}
+and (according to Rules 5 and 6 of the IA-DL method) the next iteration falls into the interval $[x_{j-1}, x_j]$, which contradicts our assumption regarding the non-limit nature of the point $\bar x$.
+The theorem is proven.
+
+\section{Numerical experiments}
 
 \section{TEST}
-
-There are various bibliography styles available.
-You can select the style of your choice in the preamble of this document.
-These styles are Elsevier styles based on standard styles like Harvard and Vancouver.
-Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.
-
-Here are two sample references:
 

### 2020-01-21 2019/NUMTA_Barkalov/Special Issure/article.tex
Move images to main directory
add some references

@@
 \begin{keyword}
-Global optimization, Multiextremal problems, Non-convex constraints, Lipschitz constant estimates 
+Global optimization, Multiextremal problems, Non-convex constraints, Lipschitz constant estimates, Dimensionality reduction, Numerical methods
 \end{keyword}
@@
 Note that these global optimization algorithms are, as a rule, oriented towards solving either  unconstrained optimization problems
-	\cite{Zilinskas, Sergeyev, Pinter, Jones}
+	\cite{Paulavicius2014, Sergeyev2015, Pinter1996, Jones1993, Gablonsky2001}
 or problems with constraints of a simple structure \cite{Vaz2009, Paulavicius2016}.
@@
 	Within the framework of our research, an original approach to the minimization of the multiextremal functions with non-convex constraints called the index method of the accounting for the constraints
-	\cite{Strongin} 
+	\cite{Strongin2000,Pugliese,Barkalov2002,Strongin2003} 
 was used.
@@
 \begin{figure}[h]
-	\center{\includegraphics[width=0.7\linewidth]{img/epsilon_reserved_solution}}
+	\center{\includegraphics[width=0.7\linewidth]{epsilon_reserved_solution}}
 	\caption{$\varepsilon$-reserved solution}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=0.8\linewidth]{img/evolvent_N=2_m=4} \\ a)}
+		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=4} \\ (a)}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=0.8\linewidth]{img/evolvent_N=2_m=5} \\ b)}
+		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=5} \\ (b)}
 	\end{minipage}
@@
 $$
-Otherwise, execute a trial at the poin
+Otherwise, execute a trial at the point
 $$
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{img/stronginC3_IA} \\ a)}
+		\center{\includegraphics[width=\linewidth]{stronginC3_IA} \\ a)}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{img/stronginC3_IA-DL} \\ b)}
+		\center{\includegraphics[width=\linewidth]{stronginC3_IA-DL} \\ b)}
 	\end{minipage}

### 2020-01-21 2019/NUMTA_Barkalov/Special Issure/article.tex
fix one misprint

@@
 This situation is typical for the optimal design problems when some conditions for the functioning of the modeled object are not met and its other characteristics appear to be undefined.
-In this approach, the first violation of any constraint indicating that the point ybeing considered is no longer feasible terminates further computations at the point $y$.
+In this approach, the first violation of any constraint indicating that the point $y$ being considered is no longer feasible terminates further computations at the point $y$.
 

### 2020-01-21 2019/NUMTA_Barkalov/Special Issure/article.tex
add acknowledgments

@@
 
-	The dimensionality reduction scheme considered here matches a multidimensional problem with Lipschitzian minimized function and Lipschitzian constraints to a one-dimensional problem, in which the corresponding functions satisfy the uniform H\"{o}lder condition (see \cite{Strongin2000, Sergeyev2013}), i.e.
+	The dimensionality reduction scheme considered here matches a multidimensional problem with Lipschitzian objective function and Lipschitzian constraints to a one-dimensional problem, in which the corresponding functions satisfy the uniform H\"{o}lder condition (see \cite{Strongin2000, Sergeyev2013}), i.e.
 $$
@@
 
+\section*{Acknowledgments}
+This research was supported by the Russian Science Foundation, project No. 16-11-10150.
+
 \section{TEST}

### 2020-01-21 2019/NUMTA_Barkalov/Special Issure/article.tex
Add illustrations and several tables

@@
 \section{Numerical experiments}
+\begin{figure}[h!]
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=\linewidth]{Simple_59_IA} \\ a)}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=\linewidth]{Simple_59_IA-DL} \\ b)}
+	\end{minipage}
+	\caption{Solving a Simple problem by IA (a) and IA-DL (b) methods}
+	\label{fig:examples_simple}
+\end{figure}
+
+\begin{figure}[h!]
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=\linewidth]{Hard_7_IA} \\ a)}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=\linewidth]{Hard_7_IA-DL} \\ b)}
+	\end{minipage}
+	\caption{Solving a Hard problem by IA (a) and IA-DL (b) methods}
+	\label{fig:examples_simple}
+\end{figure}
+
+\begin{table}[h!]
+	\centering
+		\begin{tabular}{ccccccccccccc}
+			\hline
+			& \multicolumn{2}{c}{$N=2$} & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$} \\
+			\cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12}
+			& \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} \\
+			\hline
+			IA-DL, $r_{\nu}^{loc}=1.4$ & 316 & 766 & & 7403 & 9636 & & 44304 & 60904 & & 296357 & 763711 \\
+			\hline
+			IA-DL, $r_{\nu}^{loc}=2.0$ & 350 & 830 & & 8990(1) & 12020 & & 47133 & 57473 & & 360497 & 743874(1) \\
+			\hline
+			IA & 439 & 1074 & & 11804(1) & 20327 & & 42837 & 87310(2) & & 404660(2) & 1084990(2) \\
+			\hline
+		\end{tabular}
+	\caption{Average number of iterations}
+	\label{tab:average_number_of_iterations}
+\end{table}
+
+\begin{table}[h!]
+	\centering
+		\begin{tabular}{ccccc}
+			\hline
+			& $N=2$ & $N=3$ & $N=4$ & $N=5$ \\
+			\hline
+			IA-DL, $r_{\nu}^{loc}=1.4$ & 316 & 7403 & 44304 & 296357 \\
+			\hline
+			IA-DL, $r_{\nu}^{loc}=2.0$ & 350 & 8990(1) & 47133 & 360497 \\
+			\hline
+			IA & 439 & 11804(1) & 42837 & 404660(2) \\
+			\hline
+		\end{tabular}
+	\caption{Average number of iterations Simple}
+	\label{tab:average_number_of_iterations_simple}
+\end{table}
+
+\begin{table}[h!]
+	\centering
+		\begin{tabular}{ccccc}
+			\hline
+			& $N=2$ & $N=3$ & $N=4$ & $N=5$ \\
+			\hline
+			IA-DL, $r_{\nu}^{loc}=1.4$ & 766 & 9636 & 60904 & 763711 \\
+			\hline
+			IA-DL, $r_{\nu}^{loc}=2.0$ & 830 & 12020 & 57473 & 743874(1) \\
+			\hline
+			IA & 1074 & 20327 & 87310(2) & 1084990(2) \\
+			\hline
+		\end{tabular}
+	\caption{Average number of iterations Hard}
+	\label{tab:average_number_of_iterations_hard}
+\end{table}
+
+\begin{figure}[h!]
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_3} \\ a)}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_3} \\ b)}
+	\end{minipage}
+	\caption{Operational characteristics for Simple (a) and Hard (b) problems, $N=3$}
+	\label{fig:operational_characteristics_N=3}
+\end{figure}
+
+\begin{figure}[h!]
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_5} \\ a)}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_5} \\ b)}
+	\end{minipage}
+	\caption{Operational characteristics for Simple (a) and Hard (b) problems, $N=5$}
+	\label{fig:operational_characteristics_N=5}
+\end{figure}
+
+\begin{figure}[h!]
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=\linewidth]{Decision_rule_gklssC_4} \\ a)}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[h]{0.48\linewidth}
+		\center{\includegraphics[width=\linewidth]{Decision_rule_gklshC_4} \\ b)}
+	\end{minipage}
+	\caption{Frequency of using various estimates of the Lipschitz constant for Simple (a) and Hard (b) problems, $N=4$}
+	\label{fig:frequency_N=4}
+\end{figure}
 
@@
 \section{TEST}
-
 \cite{Evtushenko1971, Piyavskii1972, Shubert1972, Strongin1970, Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Pinter1996, Jones2009, Wood1991, Meewella1988, Mladineo1986, Vaz2009, Stripinis2019, Paulavicius2016, Pillo2012, Pillo2016, Barkalov2017_1, Barkalov2017_2, Sergeyev2006, Zilinskas2008, Sovrasov2019, Kvasov2003, Sergeyev2010,Sergeyev2016, Horst1996, Gablonsky2001, Jones1993, Gaviano2003, Barkalov2018, Paulavicius2014, Sergeyev2015, Strongin2018, Gergel2017_2, Gergel2019}.

### 2020-01-24 2019/NUMTA_Barkalov/Special Issure/article.tex
Added Numerical experiments

@@
 \section{Numerical experiments}
+	To demonstrate the efficiency of the proposed algorithm, let us consider the results of solving several series of multidimensional constrained global optimization problems.
+In order to construct the problems, the functions generated by the GKLS generator \cite{Gaviano2003} were used.
+This generator of multiextremal functions is often used for investigating global optimization algorithms \cite{Barkalov2018, Paulavicius2014, Sergeyev2015}.
+According to \cite{Gaviano2003}, we used simpler and more complex functions, which are  further denoted as “Simple” and “Hard”, respectively.
+The complexity of the functions is determined, in particular, by the radius of the global minimizer's attraction region and by the distance between global and local minimizers.
+
+	In \cite{Gergel2017_2, Gergel2019}, an approach was proposed that allows constructing the constrained global optimization problems using certain classes of multiextremal functions (in particular, the GKLS functions), in which:
+\begin{itemize}
+  \item the number of constraints can be varied;
+  \item the volume of the feasible subdomain $Q_{m+1}$ relative to the whole search domain $D$ is defined;
+	\item The global minimizer is known in advance and can be located at the boundary of the feasible domain;
+	\item The number of constraints active in the global minimizer can be predefined.
+\end{itemize}
+
+	In the experiments that we have conducted, the number of constraints in the problems was equal to 2, and one of these constraints was active at the point of the global minimum.
+For example, the contour plots of the objective functions and constraints for $N=2$ are presented in Figs.
+\ref{fig:examples_simple} and \ref{fig:examples_hard}; the feasible domains are highlighted by green.
+It can be seen that the feasible domains in these problems are non-convex and non-connected, and the global minimizers are located at the boundaries of the feasible domains.
+Also, one can see that the problems of the "Hard" class have, in general, a more complex structure: the global minimizers are located at the boundaries of the feasible domains of smaller areas whereas in the problems of the “Simple” class the optimizers are located at the boundaries of the subdomains of greater areas.
+
 \begin{figure}[h!]
@@
 	\end{minipage}
-	\caption{Solving a Simple problem by IA (a) and IA-DL (b) methods}
+	\caption{Solving a problem of the “Simple” class by IA (a) and IA-DL (b) methods}
 	\label{fig:examples_simple}
@@
 	\end{minipage}
-	\caption{Solving a Hard problem by IA (a) and IA-DL (b) methods}
-	\label{fig:examples_simple}
+	\caption{Solving a problem of the “Hard” class by IA (a) and IA-DL (b) methods}
+	\label{fig:examples_hard}
 \end{figure}
 
-\begin{table}[h!]
-	\centering
-		\begin{tabular}{ccccccccccccc}
-			\hline
-			& \multicolumn{2}{c}{$N=2$} & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$} \\
-			\cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12}
-			& \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} \\
-			\hline
-			IA-DL, $r_{\nu}^{loc}=1.4$ & 316 & 766 & & 7403 & 9636 & & 44304 & 60904 & & 296357 & 763711 \\
-			\hline
-			IA-DL, $r_{\nu}^{loc}=2.0$ & 350 & 830 & & 8990(1) & 12020 & & 47133 & 57473 & & 360497 & 743874(1) \\
-			\hline
-			IA & 439 & 1074 & & 11804(1) & 20327 & & 42837 & 87310(2) & & 404660(2) & 1084990(2) \\
-			\hline
-		\end{tabular}
-	\caption{Average number of iterations}
-	\label{tab:average_number_of_iterations}
-\end{table}
+	Also, the points of the trials performed by (a) the original index algorithm (IA) and by (b) the method with dual estimate of the Lipschitz constant (IA-DL) are marked in Figs.
+\ref{fig:examples_simple} and \ref{fig:examples_hard}.
+To solve the problems shown in Figs.
+\ref{fig:examples_simple} and \ref{fig:examples_hard}, the IA method required 643 and 1166 iterations whereas the IA-DL method solved similar problems for 373 and 712 iterations, respectively.
+
+	The reduction of the number of trials was also confirmed by solving several series of  problems composed of the functions of the “Simple” and “Hard” classes.
+Each series consisted of 100 problems and was solved by the IA and IA-DL methods.
+The global minimizer $y^{\ast}$ was considered to be found, if the algorithm generated a trial point $y^k$ in the vicinity of the global minimum i.
+e.
+$\| y^{\ast} - y^k \| < \varepsilon \|b - a\|$ where $a$ and $b$ are the boundaries of the search domain $D$.
+To reduce the dimensionality, the evolvent constructed with the parameter $m=10$ was used in both algorithms.
+The relative accuracy of the search for the solution at $N = 2, 3$ was $\epsilon=0.01$ and for the solution at $N=4,5$  it was $\epsilon=0.03$.
+The reservation parameters $\epsilon_{\nu}=\mu_{\nu} \delta,1 \leq \nu \leq 2$ were set according to formula (\ref{epsilon_nu}) with $\delta=0.01$.
+The maximum number of iterations at $N \leq 4$ was set as $K_{max} = 10^6$, and at $N=5$ as $K_{max}=7 \cdot 10^6$.
+
+	Since all functions involved in the statement of each particular problem belong to the same class, equal parameters $r_{\nu}=4.1$ were set for the problems of the “Simple” class and $r_{\nu}=6.2$ for the problems of the “Hard” class.
+When solving the problems of these classes by the IA-DL method, the values of the parameters $r_{\nu}$ given above were selected for the upper estimate of the Lipschitz constants i.
+e.
+the values $r_{\nu}^{glob}=4.1$ and $r_{\nu}^{glob}=6.2$ were set, which were supplemented by various values of $r_{\nu}^{loc}$.
+The average number of iterations performed by the algorithms with different values of $r_{\nu}^{loc}$ is given in Table \ref{tab:average_number_of_iterations_simple} and \ref{tab:average_number_of_iterations_hard}.
+The number of unsolved problems is given in brackets.
 
@@
 		\end{tabular}
-	\caption{Average number of iterations Simple}
+	\caption{Average number of iterations performed by the IA and IA-DL methods on “Simple” class}
 	\label{tab:average_number_of_iterations_simple}
@@
 		\end{tabular}
-	\caption{Average number of iterations Hard}
+	\caption{Average number of iterations performed by the IA and IA-DL methods on “Hard” class}
 	\label{tab:average_number_of_iterations_hard}
 \end{table}
+
+	In addition to the average number of trials required to solve the problems, the results of solving the series of problems can be also presented by the function $p(k)$, which describes the number of problems solved for $k$ iterations.
+This function will be called the \textit{operational characteristic} of the algorithms.
+
+	The operational characteristics for IA and IA-DL obtained by solving the series of problems of the “Simple” and “Hard” classes with the dimensionalities $N=3$ and $N=5$ are presented in Figs.
+\ref{fig:operational_characteristics_N=3} and \ref{fig:operational_characteristics_N=5}, respectively.
+The values of parameters $r_{\nu}$ used to evaluate the Lipschitz constant are given in the figures.
 
@@
 	\end{minipage}
-	\caption{Operational characteristics for Simple (a) and Hard (b) problems, $N=3$}
+	\caption{Operational characteristics for “Simple” (a) and “Hard” (b) problem classes, $N = 3$}
 	\label{fig:operational_characteristics_N=3}
@@
 	\end{minipage}
-	\caption{Operational characteristics for Simple (a) and Hard (b) problems, $N=5$}
+	\caption{Operational characteristics for “Simple” (a) and “Hard” (b) problem classes, $N = 5$}
 	\label{fig:operational_characteristics_N=5}
 \end{figure}
+
+	The lower (solid) curves in Figs.
+\ref{fig:operational_characteristics_N=3} and \ref{fig:operational_characteristics_N=5} characterize IA whereas the upper (dashed) lines characterize IA-DL.
+This arrangement of curves shows that the algorithm with dual estimates of the Lipschitz constant  provides, on average, a faster solution to the problems of the series than the algorithm using a single estimate of the Lipschitz constant.
+
+	The frequency of the use of the upper and lower estimates of the Lipschitz constant in the rules of the algorithm that corresponds to the choice of $r_{\nu}=r_{\nu}^{loc}$ or $r_{\nu}=r_{\nu}^{glob}$ according to Step 6 of the algorithm is an interesting indicator.
+This frequency can be characterized by the average values $K_{glob}$ and $K_{loc}$ of the use of the respective (global or local) characteristic.
+In order to present this frequency, let us plot the number of trials normalized from 1 to 100 on the abscissa axis.
+Actually, in this case the abscissa axis will reflect the progress in solving the problem (in percentage) where 1 corresponds to the beginning and 100 -- to the end of the process of the search for the solution.
+The indicators $K_{glob}$ and $K_{loc}$ will be plotted on the ordinate axis.
+Such plots are presented in Fig.
+\ref{fig:frequency_N=4}.
 
@@
 	\end{minipage}
-	\caption{Frequency of using various estimates of the Lipschitz constant for Simple (a) and Hard (b) problems, $N=4$}
+	\caption{Frequency of the use of different estimates of the Lipschitz constant for the problems of the “Simple” class (a) and of the “Hard” class (b) for $N=4$}
 	\label{fig:frequency_N=4}
 \end{figure}
+
+	The dashed lines in Fig.
+\ref{fig:frequency_N=4} correspond to the frequency of the use of the lower estimates of the Lipschitz constant, the solid lines -- to the frequency of the use of the upper estimate.
+The arrangement of the curves shows the lower estimate of the Lipschitz constant to be used more frequently  at the beginning of the search process whereas at the final stage both estimates are used with almost equal frequency.
+These plots coincide qualitatively with the frequencies of the use of different estimates of the Lipschitz constant for other series of problems as well.
 
@@
 
-\section{TEST}
-\cite{Evtushenko1971, Piyavskii1972, Shubert1972, Strongin1970, Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Pinter1996, Jones2009, Wood1991, Meewella1988, Mladineo1986, Vaz2009, Stripinis2019, Paulavicius2016, Pillo2012, Pillo2016, Barkalov2017_1, Barkalov2017_2, Sergeyev2006, Zilinskas2008, Sovrasov2019, Kvasov2003, Sergeyev2010,Sergeyev2016, Horst1996, Gablonsky2001, Jones1993, Gaviano2003, Barkalov2018, Paulavicius2014, Sergeyev2015, Strongin2018, Gergel2017_2, Gergel2019}.
-
 \bibliography{bibliography}

### 2020-01-24 2019/NUMTA_Barkalov/Special Issure/article.tex
Added Conclusion

@@
 
+\section{Conclusion}
+	In this paper, a method for solving the Lipschitz optimization problems, which constitute an important class of the global optimization problems, was considered.
+In applied global optimization problems, the functions are usually defined as the “black boxes”.
+Consequently, the values of the Lipschitz constant for them are unknown \textit{a priori}, which makes their estimation one of the key problems when constructing the Lipschitz optimization methods.
+
+	A correct estimate of the Lipschitz constant is especially important in solving constrained problems since the feasible domain in these problems has a complex structure, which is largely determined by the values of the constants for the constraints of the problem.
+An underestimate of the true value of the constant may lead to the loss of convergence of the method to the global optimizer whereas an overestimated value results in a slow convergence to the solution of the problem.
+
+	This paper proposes an algorithm for solving global optimization problems with non-convex constraints, which uses two Lipschitz constant estimates, one of which is considerably larger than the other.
+The larger estimate ensures global convergence and the smaller one reduces the total number of trials required to find the global optimizer.
+The choice of the estimate to be used in the rules of the algorithm is performed automatically.
+The conditions of convergence for the algorithm with dual estimate of the Lipschitz constant were found.
+The numerical experiments carried out on a series of several hundred constrained global optimization test problems demonstrated the effect of reduction of the number of search trials in the index algorithm with dual estimate of the Lipschitz constant as compared to the original version of the algorithm.
+
+	Further research areas include the use of the dual estimate of the Lipschitz constant in parallel algorithms, which can be constructed with the use of various schemes for  dimensionality reduction and  for parallelization of the computations \cite{Strongin2018}.
+
 \section*{Acknowledgments}

### 2020-01-24 2019/NUMTA_Barkalov/Special Issure/article.tex
some minor corrections in references ans formulas

@@
-\documentclass[review]{elsarticle}
+\documentclass[preprint]{elsarticle}
 
@@
 	The first four steps of the algorithm repeat Step 1 -- Step 4 of the initial index algorithm from Section 3.
-	Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$ 
+	Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1,$ 
 \begin{flushleft}
@@
 The same values were used for the IA-DL method as the parameters $r_{\nu}^{glob}, 1 \leq \nu \leq 4$.
-Also, the parameters $r_{\nu}^{loc}=1.5, 1 \leq \nu \leq 4$ were used.
+Also, the parameters $r_{\nu}^{loc}=1.5, 1 \leq \nu \leq 4,$ were used.
 To solve the problem with a predefined accuracy, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method.
@@
 Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1$ the conditions of Theorem 1 are fulfilled.
-Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1, 1\leq \nu \leq m+1$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e.
+Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1, 1\leq \nu \leq m+1,$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e.
 satisfies the condition
@@
 \end{equation}
+
 \textbf{Proof}.
-
-	Because for the IA with the parameters $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1$ the conditions of Theorem 1 are fulfilled, then any limit point $\bar{y}=y(\bar{x})$ of the trial sequence generated by the IA for the problem (\ref{problem}) satisfies the conditions (\ref{phi_bar_y}).
+Because for the IA with the parameters $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem 1 are fulfilled, then any limit point $\bar{y}=y(\bar{x})$ of the trial sequence generated by the IA for the problem (\ref{problem}) satisfies the conditions (\ref{phi_bar_y}).
 

### 2020-01-25 2020/PAVT_Sovrasov/pavt_paper.tex
PAVT: add one more author
PAVT: fix style issues

@@
 \usepackage{subfig}
-\usepackage{caption}
-%\captionsetup[table]{skip=9pt}
-%\captionsetup{labelfont=bf}
+\usepackage[labelfont=bf,labelsep=period]{caption}
 
@@
 
-\authors{В.В.~Соврасов}
+\authors{В.В.~Соврасов, К.A.~Баркалов}
 \organizations{Нижегородский государственный университет им. Н.И.~Лобачевского}
@@
 в процессе работы алгоритма погрешности численного решения во всех задачах убывают примерно с одинаковой скоростью.
-Подобные серии задач возникают, если задача глобальной оптимизации имеет дискретный параметр или при решении задачи многокритериальной оптимизации методом свёртки критериев.
+Подобные серии задач возникают, если задача глобальной оптимизации имеет дискретный параметр или при решении задачи многокритериальной оптимизации методом свертки критериев.
 Рассматриваемый алгоритм использует отображения типа кривой Пеано для редукции многомерных задач оптимизации к одномерным.
@@
 Решение серии таких задач при ограниченных вычислительных
-ресурсах является ещё более сложной задачей: помимо поиска глобального экстремума необходимо
+ресурсах является еще более сложной задачей: помимо поиска глобального экстремума необходимо
 распределять вычислительные ресурсы так, чтобы сразу во всех решаемых задачах положение глобального
@@
 параллельно порциями по \(p,\:p<<q\) задач, где \(p\) --- количество параллельных вычислительных устройств.
-Такой подход ведёт к тому, что в каждый момент времени до окончания вычислений
+Такой подход ведет к тому, что в каждый момент времени до окончания вычислений
 остаются задачи, в которых оценка глобального оптимума не получена вообще, в то время, как в задачах из начала
-списка оптимум может быть оценён даже с избыточной точностью.
+списка оптимум может быть оценен даже с избыточной точностью.
 
@@
 параллельного метода глобальной оптимизации для одновременного решения множества задач \cite{BarkalovStrongin2018} на
-случай задач с нелинейными ограничениями. Для учёта ограничений используется индексная схема \cite{Strongin2000},
+случай задач с нелинейными ограничениями. Для учета ограничений используется индексная схема \cite{Strongin2000},
 позволяющая работать с частично вычислимыми целевым функциями и обладающая экономичностью,
@@
 на множестве задач, возникающем при решении задачи многокритериальной оптимизации
-с нелинейными ограничениями методом свёртки критериев \cite{Ehrgott2005}.
+с нелинейными ограничениями методом свертки критериев \cite{Ehrgott2005}.
 
@@
 Отображение вида (\ref{cube}) позволяет свести задачу в многомерном пространстве к решению
-одномерной ценой ухудшения её свойств. В частности, одномерная функция \(\varphi(y(x))\)
+одномерной ценой ухудшения ее свойств. В частности, одномерная функция \(\varphi(y(x))\)
 является не Липшицевой, а Гёльдеровой:
@@
 точек \(x_k\), в которых вычисляются значения минимизируемой функции или ограничений \(z_k = g_s(x_k)\).
-Для учёта последних используется индексная схема \cite{Strongin2000}. Пусть \(Q_0=[0;1]\). Ограничение, имеющее номер
+Для учета последних используется индексная схема \cite{Strongin2000}. Пусть \(Q_0=[0;1]\). Ограничение, имеющее номер
  \(j\), выполняется во всех точках области
@@
 
-Вместе с системой GCGen распространяются примеры её использования и построения
+Вместе с системой GCGen распространяются примеры ее использования и построения
 наборов задач, каждая из которых состоит из целевой функции и двух ограничений,
-порождённых генератором \(F_{GR}\) \cite{grishaginClass} или GKLS \cite{Gaviano2003}.
+порожденных генератором \(F_{GR}\) \cite{grishaginClass} или GKLS \cite{Gaviano2003}.
 
@@
     \centering
-    \subfloat[]{
+    \subfloat[Решение внутри допустимой области]{
     {\includegraphics[width=.5\textwidth]{images/2.png}}}
-    \subfloat[]{
+    \subfloat[Решение на границе допустимой области]{
     {\includegraphics[width=.5\textwidth]{images/4.png}}}
@@
 
-Будем считать, что тестовая задача решена, если метод оптимизации провёл очередное испытание \(y^k\) в
+Будем считать, что тестовая задача решена, если метод оптимизации провел очередное испытание \(y^k\) в
 \(\delta\)-окрестности глобального минимума \(y^*\), т.е. $\left\|y^k-
 y^*\right\|\leqslant \delta = 0.01\left\|b-a\right\|$, где \(a\) и \(b\) --- левая и правая границы гиперкуба из (\ref{eq:task}).
-Если указанное соотношение не выполнено до истечения лимита на количество испытаний, то задача считается нерешённой.
+Если указанное соотношение не выполнено до истечения лимита на количество испытаний, то задача считается нерешенной.
 
 %В качестве характеристик метода оптимизации на каждом из классов будем рассматривать среднее число
-%испытаний, затраченное для решения одной задачи, и количество решённых задач. Чем меньше число испытаний, тем быстрее метод сходится
-%к решению, а значит и меньше обращается к потенциально трудоёмкой процедуре вычислений целевой функции.
+%испытаний, затраченное для решения одной задачи, и количество решенных задач. Чем меньше число испытаний, тем быстрее метод сходится
+%к решению, а значит и меньше обращается к потенциально трудоемкой процедуре вычислений целевой функции.
 При оценке качества метода и его реализации кроме ускорения от распараллеливания и времени выполнения также будем принимать во внимание среднее максимальное расстояния (в смысле \(l_{\inf}\)-нормы) текущей оценки оптимума до его реального положения,
@@
 для решения множества задач представлены в таблице \ref{tab:speedup}. Для всех двухмерных классов задач параметр \(r=4.7\).
-В случае трёхмерных задач \(r=4.7,\: \varepsilon_\nu=0.1\).
+В случае трехмерных задач \(r=4.7,\: \varepsilon_\nu=0.1\).
 Во всех экспериментах в целевые функции и ограничения была внесена дополнительная вычислительная нагрузка так,
@@
 
-Из таблицы видно, что ускорение по итерациям \(S_i\) растёт линейно с увеличением числа потоков \(p\),
+Из таблицы видно, что ускорение по итерациям \(S_i\) растет линейно с увеличением числа потоков \(p\),
 в то время, как ускорение по времени \(S_p\) увеличивается не так быстро, что говорит о неидеальной
 реализации алгоритма. Увеличить реальное ускорение, верхней границей для которого является
-\(S_i\), возможно путём оптимизации взаимодействий между копиями ИАГП и это планируется сделать в ходе будущей работы.
+\(S_i\), возможно путем оптимизации взаимодействий между копиями ИАГП и это планируется сделать в ходе будущей работы.
 
@@
 средних и максимальных расстояний от реальных оптимумов до их текущих оценок при решении
-серии из задач, порождённых двумя разными генераторами, по отдельности (сплошная кривая) и совместно (пунктирная кривая).
+серии из задач, порожденных двумя разными генераторами, по отдельности (сплошная кривая) и совместно (пунктирная кривая).
 Не смотря на значительную разницу в структуре задач, модифицированный ИАГП
@@
     \centering
-    \subfloat[]{
+    \subfloat[\(D_{max}\)]{
     \label{fig:max_dev} {\includegraphics[width=.5\textwidth]{images/mixed_2d_max.pdf}}}
-    \subfloat[]{
+    \subfloat[\(D_{avg}\)]{
     \label{fig:avg_dev} {\includegraphics[width=.5\textwidth]{images/mixed_2d_avg.pdf}}}
-    \caption{Динамика величин \(D_{avg}\) и \(D_{max}\) в процессе решения множества задач}
+    \caption{Динамика величин \(D_{avg}\) и \(D_{max}\) в процессе решения множества двухмерных задач,
+    порождённых двумя разными генераторами GKLS и \(F_{GR}\)}
     \label{fig:devs_mixed}
@@
 
-Будем использовать свёртку Гермейера для скаляризации задачи (\ref{eq:mco_probem}).
-После свёртки скалярная целевая функция имеет вид:
+Будем использовать свертку Гермейера для скаляризации задачи (\ref{eq:mco_probem}).
+После свертки скалярная целевая функция имеет вид:
 \begin{equation}
@@
 где \(\lambda_1,\lambda_2\in[0,1],\: \lambda_1+\lambda_2=1\). Перебирая все возможные
-коэффициенты свёртки, можно найти всё множество парето-оптимальных решений в
+коэффициенты свертки, можно найти все множество парето-оптимальных решений в
 задаче (\ref{eq:mco_probem}). Для численного построения множества Парето выберем
@@
   25 испытаний. Таким образом, вычислительные ресурсы равномерно распределены между задачами;
-  \item все задачи решаются одновременно с помощью обобщённого ИАГП с установленным лимитом в
+  \item все задачи решаются одновременно с помощью обобщенного ИАГП с установленным лимитом в
   2500 испытаний.
@@
     \centering
-    \subfloat[ИАГП, раздеьное решение задач]{
+    \subfloat[ИАГП, раздельное решение задач]{
     \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{images/single_mco.pdf}}}
@@
 показатели параллельного ускорения по времени. Также планируется реализовать версию
-рассматриваемого алгоритма, работающего на распределённой памяти по схеме, описанной в \cite{BarkalovLebedev2017_2}.
+рассматриваемого алгоритма, работающего на распределенной памяти по схеме, описанной в \cite{BarkalovLebedev2017_2}.
 
@@
 %\bibitem{eremin}
-%Ерёмин~И.И. Фейеровские методы для задач выпуклой и линейной оптимизации.
+%Еремин~И.И. Фейеровские методы для задач выпуклой и линейной оптимизации.
 %Челябинск: Изд-во ЮУрГУ, 2009. 200~с.

### 2020-01-27 2019/NUMTA_Barkalov/Special Issure/article.tex
a lot of minor corrections

@@
 The upper estimate provides the global convergence whereas the lower one reduces the number of trials necessary to find the global optimizer with the required accuracy.
-In this case,  the considered algorithm for solving the constrained problems doesn’t use the ideas of the penalty function method; each constraint of the problem is accounted for separately.
+The considered algorithm for solving the constrained problems doesn’t use the ideas of the penalty function method; each constraint of the problem is accounted for separately.
 The convergence conditions of the proposed algorithm are formulated in the corresponding theorem.
@@
 As a rule, problem functions are defined by a program code, i.e.
-they are “black-box” functions, for which the computing of the values can be a computationally costly operation since in applied problems it will require  numerical modeling.
-Any  possibility to estimate reliably the global optimum in the multiextremal problems with “black-box” functions is based fundamentally on the auxiliary assumptions relating the possible values of the problem functions to their known values at the points of the trials already performed.
+they are ``black-box'' functions, for which the computing of the values can be a time-consuming operation since in applied problems it will require  numerical modeling.
+Any  possibility to estimate reliably the global optimum in the multiextremal problems with ``black-box'' functions is based fundamentally on the auxiliary assumptions relating the possible values of the problem functions to their known values at the points of the trials already performed.
 
@@
 This approach is based on separate accounting for each constraint of the problem and does not involve the use of the penalty functions.
-According to the rules of the index method, every iteration called trial at the corresponding point of the search domain includes a sequential checking of the feasibility of the problem constraints at this point.
+According to the rules of the index method, every iteration called \textit{trial} at the corresponding point of the search domain includes a sequential checking of the feasibility of the problem constraints at this point.
 When the first violation of any constraint is found, the trial is terminated (the values of the remaining problem functions are not computed at this point) and the transition to the next iteration point is initiated.
@@
 	Note that the values of the Lipschitz constant for the problem functions are usually not known a priori, which  makes their estimation one of the key problems in the construction of the Lipschitzian optimization methods.
-The methods utilizing the values of the Lipschitz constant predefined a priori (see, for example, \cite{Piyavskii1972, Shubert1972, Wood1991, Meewella1988, Mladineo1986}) are important in the theoretical aspect, but it is difficult to apply such methods in applied problems (in which the information of the Lipschitz constant values is absent).
+The methods utilizing the values of the Lipschitz constant predefined a priori (see, for example, \cite{Piyavskii1972, Shubert1972, Wood1991, Meewella1988, Mladineo1986}) are important in the theoretical aspect, but it is difficult to use such methods in applied problems (in which the information of the Lipschitz constant values is absent).
 
@@
 
-		Several typical methods for adaptive estimation of the Lipschitz constant according to the results of the performed search trial are known:
+		Several typical methods for adaptive estimation of the Lipschitz constant according to the results of the performed search trials are known:
 \begin{itemize}[\labelitemii]
-  \item global estimate of the constant $L$ in the whole search domain $D$ \cite{Horst1996, Pinter1996, Strongin2000}.
+  \item global estimate of the constant $L$ in the whole search domain $D$ \cite{Strongin2000,Pinter1996,Horst1996}.
   \item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ \cite{Kvasov2003, Sergeyev2010, Sergeyev2016}.
-	\item choice of the estimates of the constant $L$ from some set of possible values \cite{Gablonsky2001, Jones1993, Jones2009, Sergeyev2006}.
+	\item choice of the estimates of the constant $L$ from some set of possible values \cite{Jones2009,Jones1993, Gablonsky2001,Sergeyev2006}.
 \end{itemize}
@@
 \end{equation}
-with \textit{a priori} unknown constant $L_j, 1\leq j \leq m+1$, and may be multiextremal.
+with a priori unknown constants $L_j, 1\leq j \leq m+1$, and may be multiextremal.
 
@@
 	\center{\includegraphics[width=0.7\linewidth]{epsilon_reserved_solution}}
-	\caption{$\varepsilon$-reserved solution}
+	\caption{$\varepsilon$-reserved solution.}
 	\label{fig:eps_reserved_solution}
@@
 
-	We will consider the initial problem (\ref{problem}) assuming that the problem functions $g_{j}(y), 1 \leq j \leq m+1$ may be defined only partly.
-It means that they are defined and calculable only in the sets
+	We will consider the initial problem (\ref{problem}) assuming that the problem functions $g_{j}(y), 1 \leq j \leq m+1,$ may be defined only partly.
+It means that they are defined and computable only in the sets
 \begin{equation}\label{D_sets}
@@
 \begin{equation}\label{one_dimensional_problem}
-	\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\},
+	\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
 \end{equation}
 The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000, Sergeyev2013}.
-Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$ where the number $m>1$ is a parameter of the construction (the density of the evolvent).
-In this case, the construction density of the evolvent is determined by the accuracy of the possible practical implementation of the problem solution found.
+Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$, where the number $m>1$ is a parameter (the \textit{density} of the evolvent).
+The evolvent density is determined by the accuracy of the possible practical implementation of the problem solution found.
 For  illustration, Fig.
-\ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] are shown for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
+\ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
 \begin{figure}[h]
@@
 
-	The dimensionality reduction scheme considered here matches a multidimensional problem with Lipschitzian objective function and Lipschitzian constraints to a one-dimensional problem, in which the corresponding functions satisfy the uniform H\"{o}lder condition (see \cite{Strongin2000, Sergeyev2013}), i.e.
+	The dimensionality reduction scheme considered here matches a multidimensional problem with Lipschitzian objective function and Lipschitzian constraints to a one-dimensional problem, in which the corresponding functions satisfy the uniform H\"{o}lder condition (see \cite{Strongin2000, Sergeyev2013})
 $$
@@
 
-	According to (\ref{D_sets}) and (\ref{one_dimensional_problem}), the reduced one-dimensional functions $g_j (y(x)), 1 \leq j \leq m+1$ will also be defined and calculable in the corresponding intervals
+	According to (\ref{D_sets}) and (\ref{one_dimensional_problem}), the reduced one-dimensional functions $g_j (y(x)), 1 \leq j \leq m+1,$ will also be defined and computable in the corresponding intervals
 \begin{equation}\label{Q_intervals}
@@
 \end{equation}
-These relations allow introducing a classification of the points $x \in [0,1]$ according to the number $\nu = \nu(x)$ of the constraints calculable at the given point.
+These relations allow introducing a classification of the points $x \in [0,1]$ according to the number $\nu = \nu(x)$ of the constraints computable at the given point.
 We will call this number $\nu = \nu(x)$ the \textit{index} of the point $x$.
@@
 
-	In the case when the point $y^k$ is a feasible one, i.e.
-when $y(x^k) \in Q_(m+1)$, the trial involves checking all constraints and computing the objective function.
+	In the case when the point $y^k = y(x^k)$ is a feasible one, i.e.
+when $x^k \in Q_{m+1}$, the trial involves checking all constraints and computing the objective function.
 Here, the value of the index is accepted to be equal to the quantity $\nu = m+1$.
@@
 
-	The value $\psi(x)$ at $\nu(x) \leq m$ matches the value of the left-hand part of the first constraint violated at this point $g_{\nu}(y(x))$ divided by the constant $K_{\nu}$.
+	The value $\psi(x)$ at $\nu(x) \leq m$ matches the value of the left-hand side of the first constraint violated at this point $g_{\nu}(y(x))$ divided by the constant $K_{\nu}$.
 In the case when $\nu(x) = m+1$
@@
 	The index-based global search algorithm under consideration for solving the problem (\ref{problem}) (here it is given according to \cite{Strongin2000}) involves the construction of a sequence of points $x^k \in (0,1)$, at which the search trials are conducted with the results of the type (\ref{trial_result}).
-The set of triplets $\{x^k, z^k=g_{\nu}\big(y(x^k)\big), \nu=\nu(x^k)\}, 1 \leq k \leq n$  makes up the search information accumulated by the method after performing $n$ steps.
+The set of triplets $\{x^k, z^k=g_{\nu}\big(y(x^k)\big), \nu=\nu(x^k)\}, 1 \leq k \leq n,$  makes up the search information accumulated by the method after performing $n$ steps.
 The operation of the index-based global search algorithm is defined by the following rules.
@@
 $$
-including the numbers of all points $x_i, 1 \leq i \leq k$ with the same values of $\nu$.
+including the numbers of all points $x_i, 1 \leq i \leq k,$ with the same values of $\nu$.
 The end points $x_0=0$ and $x_{k+1}=1$ are interpreted as the ones having the indices equal to zero.
@@
 \begin{equation}\label{current_lower_bounds}
-	\mu_{\nu} = \max\left\{ \frac{\left|z_i-z_j\right|}{ (x_i - x_j)^{1/N} }, \; i,j \in I_\nu, \; i>j \right\},
+	\mu_{\nu} = \max\left\{ \frac{\left|z_i-z_j\right|}{ (x_i - x_j)^{1/N} }, \; i,j \in I_\nu, \; i>j \right\}
 \end{equation}
@@
   \item Problem (\ref{problem}) has an $\varepsilon$-reserved solution $y_{\varepsilon}$ from (\ref{epsilon_reserved_solution}).
-  \item Functions $g_j(y), 1 \leq j \leq m+1$ admit Lipschitzian (with the constants $L_j$) extensions $G_{j}(y)$ over the whole feasible domain $D$ from (\ref{D}), i.e.
+  \item Functions $g_j(y), 1 \leq j \leq m+1,$ admit Lipschitzian (with the constants $L_j$) extensions $G_{j}(y)$ over the whole feasible domain $D$ from (\ref{D}), i.e.
 $$
@@
 where the sets $Q_j$ are from (\ref{Q_intervals}), and $y(x)$ is the curve from (\ref{one_dimensional_problem}).
-	\item The parameters of the method $\epsilon_{\nu}, 1 \leq \nu \leq m$ used in the rule (\ref{z_estimates}) are the corresponding components of the reserve vector $\epsilon_{R}$ from (\ref{epsilon_reserved_solution}).
-	\item For a sufficiently large iteration number $q$, the values $\mu_{\nu}$ from (\ref{current_lower_bounds}) satisfy the inequalities
+	\item The parameters of the method $\epsilon_{\nu}, 1 \leq \nu \leq m,$ used in the rule (\ref{z_estimates}) are the corresponding components of the reserve vector $\epsilon_{R}$ from (\ref{epsilon_reserved_solution}).
+	\item For a sufficiently large iteration number $k$, the values $\mu_{\nu}$ from (\ref{current_lower_bounds}) satisfy the inequalities
 		\begin{equation}\label{theorem_inequalities}
-			r_{\nu}\mu_{\nu} > 2^{3-\frac{1}{N}}L_{\nu}\sqrt{N+3}, \; 1 \leq \nu \leq m+1, 
+			r_{\nu}\mu_{\nu} > 2^{3-\frac{1}{N}}L_{\nu}\sqrt{N+3}, \; 1 \leq \nu \leq m+1. 
 		\end{equation}
@@
 $$
-	\varphi(\bar y) = \inf\left\{\varphi(y^q):g_j(y^q) \leq 0, 1 \leq j \leq m, q = 1,2,\ldots \right\} \leq \varphi(y_{\epsilon}).
+	\varphi(\bar y) = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \leq \varphi(y_{\epsilon}).
 $$
@@
 The software implementation of this scheme may be made more frugal.
-For example, according to Step 7, the point $x^{k+1}$ of the scheduled trial will be an internal point of the interval $(x_{t-1}, x_t)$, i.e.
+For example, according to Step 7, the point $x^{k+1}$ of the next trial will be an internal point of the interval $(x_{t-1}, x_t)$, i.e.
 $x^{k+1} \in (x_{t-1}, x_t)$.
@@
 \end{equation}
-where $\mu_{\nu}$ are the current estimates of the H\"{o}lder constant from (\ref{current_lower_bounds}) and $\delta>0$ is a parameter.
+where $\mu_{\nu}$ are the current estimates from (\ref{current_lower_bounds}) and $\delta>0$ is a parameter.
 This approach to choosing the reserve vector components allows one to take into account the properties of different problem functions characterized by their Lipschitz constants.
@@
 \section{Index Algorithm with Dual Lipschitz Constant Estimates}
-	The index global search algorithm described in the previous section is intended for solving  the multiextremal problems, in which the objective function and the functions of constraints satisfy the Lipschitz condition.
-In this case, \textit{a priori} estimates of the constant values are not required for the algorithm to work.
+	The index global search algorithm described in the previous section (hereafter referred to as IA) is intended for solving  the multiextremal problems, in which the objective function and the constraints satisfy the Lipschitz condition.
+In this case, a priori estimates of the constant values are not required for the algorithm to work.
 These estimates are computed in the course of global search, based on the available search information.
@@
 	According to Theorem 1 from Section 3, the sequence of trial points $\{y^k\}$ will converge to a global minimizer $y^\ast$ if the condition (\ref{theorem_inequalities}) is satisfied.
-So far, an appropriate choice of the parameters $r_{\nu}$ from (\ref{R_2}) allows using the values
+So far, an appropriate choice of the parameters $r_{\nu}$ allows using the values
 \begin{equation}\label{estimates_Lipschitz_constants}
-	\frac{r_{\nu}\mu_{\nu}}{2^{3-\frac{1}{N}}L_{\nu}\sqrt{N+3}}, \; 1 \leq \nu \leq m+1, 
+	\frac{r_{\nu}\mu_{\nu}}{2^{3-\frac{1}{N}}\sqrt{N+3}}, \; 1 \leq \nu \leq m+1, 
 \end{equation}
@@
 
-	Compliance with condition (\ref{estimates_Lipschitz_constants}) will be achieved when choosing a high value of the parameters $r_{\nu}$.
+	Compliance with condition (\ref{theorem_inequalities}) will be achieved when choosing a high value of the parameters $r_{\nu}$.
 However, in this case the constant will be overestimated, and the method will perform a large number of trials until the stopping condition is fulfilled.
-
 	The choice of low values of the parameters $r_{\nu}$ (which corresponds to an underestimate of the Lipschitz constant) would reduce the number of trials considerably but could destroy the convergence to the global extremum (due to possible violation of the condition (\ref{theorem_inequalities})).
@@
 	An approach, in which two estimates of the Lipschitz constant for each function are used in the rules of the algorithm, is promising.
-One of these estimates is significantly greater than the other, which corresponds to the use of two sets of parameters $r_{\nu}^{glob}$ and $r_{\nu}^{loc}$ in the algorithm where $r_{\nu}^{glob}>r_{\nu}^{loc}>1, 1 \leq \nu \leq m+1$.
+One of these estimates is significantly greater than the other, which corresponds to the use of two sets of parameters $r_{\nu}^{glob}$ and $r_{\nu}^{loc}$ in the algorithm, where $r_{\nu}^{glob}>r_{\nu}^{loc}>1, 1 \leq \nu \leq m+1$.
 When using the parameter $r_{\nu}^{loc}$ we shall deal with the smaller estimate of the Lipschitz constant, and using the parameter $r_{\nu}^{glob}$ will correspond to the larger one.
@@
 	The key idea of the novel method is computing two characteristics for each interval and balancing (based on these two) the local and global search.
-The local search provides a new trial in the vicinity of the current point $y_k^\ast$  whereas during the global search, a new trial could be performed in a subdomain located far away from $y_k^\ast$.
+The local search provides a new trial in the vicinity of the current minimum point $y_k^\ast$  whereas during the global search a new trial could be performed in a subdomain located far away from $y_k^\ast$.
 
@@
 In this case, at the final stage of the search (i.e.
-in the case when one of the points of this trial sequence falls into the vicinity of the $\varepsilon$-reserved solution $y^\ast$) there always exists an interval, at one of the boundaries of which the current minimum value of the objective function $z_{m+1}^\ast$  from (\ref{z_estimates}) is achieved.
+in the case when one of the points of this trial sequence falls into the vicinity of the $\varepsilon$-reserved solution $y_\epsilon$) there always exists an interval, at one of the boundaries of which the current minimum value of the objective function $z_{m+1}^\ast$  from (\ref{z_estimates}) is achieved.
 It is this interval that will have the highest characteristic since it is the most promising one for performing further trials.
 
-	Let us establish the following important property for the global and local characteristics of an interval containing the current optimal point.
+	Let us establish the following important property for the global and local characteristics of an interval containing the current minimum point.
 
 	\textbf{Statement}.
-Let us assume that the current minimum value $z_M^\ast$  from (\ref{z_estimates}) at the $k^{th}$ iteration is achieved at the boundary point of some interval $(x_{i-1}, x_i), i=i(k)$.
+Let us assume that the current minimum value $z_M^\ast$  from (\ref{z_estimates}) at the $k^{th}$ iteration is achieved at the boundary point of some interval $[x_{i-1}, x_i], i=i(k)$.
 Then, $R_{glob}(i) \geq R_{loc}(i)$.
@@
 \end{equation}
-  \item If $\nu(x_{i-1}) = \nu(x_i) = M$, then, taking into account the rules (\ref{current_lower_bounds}), the following inequality will hold
+  \item If $\nu(x_{i-1}) = \nu(x_i) = M$, then, taking into account (\ref{current_lower_bounds}), the following inequality will hold
 $$
-	z_i+z_{i-1}-2z_M^* = |z_i-z_{i-1}|\leq \mu_M\Delta_i
+	z_i+z_{i-1}-2z_M^* = |z_i-z_{i-1}|\leq \mu_M\Delta_i .
 $$
@@
 \nonumber \\
-	& \geq & \Delta_i + \frac{\Delta_i}{r_M^2} - 2\frac{\Delta_i}{r_M} = \Delta_i \left( 1-\frac{1}{r_M} \right)^2>0  
+	& \geq & \Delta_i + \frac{\Delta_i}{r_M^2} - 2\frac{\Delta_i}{r_M} = \Delta_i \left( 1-\frac{1}{r_M} \right)^2>0.
 	\nonumber
 \end{eqnarray}
-And because $r_{M}^{glob} > r_{M}^{loc} > 1$ then
+And because $r_{M}^{glob} > r_{M}^{loc} > 1$, then
 \begin{equation}\label{R_glob_greater_R_loc}
-	R_{glob}(i)=\Delta_i\left(1-\frac{1}{r_M^{glob}} \right)^2 > \Delta_i\left(1-\frac{1}{r_M^{loc}} \right)^2 = R_{loc}(i)
+	R_{glob}(i)=\Delta_i\left(1-\frac{1}{r_M^{glob}} \right)^2 > \Delta_i\left(1-\frac{1}{r_M^{loc}} \right)^2 = R_{loc}(i).
 \end{equation}
@@
 	The first four steps of the algorithm repeat Step 1 -- Step 4 of the initial index algorithm from Section 3.
+
 	Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1,$ 
@@
 	\begin{itemize}[\labelitemii]
-		\item Compute the global characteristics $R_{glob}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{glob}$
-		\item Compute the local characteristics $R_{loc}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{loc}$
+		\item Compute the global characteristics $R_{glob}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{glob}$.
+		\item Compute the local characteristics $R_{loc}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{loc}$.
 		\item Determine the characteristic $R(i)$ as
 \begin{equation}\label{R_max}
-	R(i) = \max{\left\{ R_{glob}(i), \rho R_{loc}(i) \right\}}
+	R(i) = \max{\left\{ R_{glob}(i), \rho R_{loc}(i) \right\}},
 \end{equation}
@@
 $$
-When computing the point $x^{k+1}$, use the value of the parameter $r_{\nu}$ fixed at Step 6.
+When computing the point $x^{k+1}$, use the value of the parameter $r_{\nu}$ fixed at Step~6.
+
 The algorithm terminates if the condition $\Delta_{t} \leq \epsilon$ is satisfied where $t$ is from (\ref{R_max_t}) and $\epsilon>0$ is the predefined accuracy.
@@
 	g_2(y_1, y_2) &=& 100 \left(1-(y_1-2)^2/1.44+(0.5y_2)^2 \right) \leq 0, \nonumber \\
-	g_3(y_1, y_2) &=& 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0
-	\nonumber
+	g_3(y_1, y_2) &=& 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0. \nonumber
 \end{eqnarray}
-The feasible domain of this problem is not connected and consists of three non-convex subdomains.
+The feasible domain of this problem is not connected and consists of three subdomains.
 The global minimum $\varphi(y^\ast)=-1.489$ is achieved at the point $y^\ast =(0.942, 0.944)$ (the values are given with the accuracy of $10^{-3}$).
@@
 Also, the parameters $r_{\nu}^{loc}=1.5, 1 \leq \nu \leq 4,$ were used.
-To solve the problem with a predefined accuracy, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method.
+To solve the problem with a predefined accuracy $\epsilon=0.002$, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method.
 The results of larger-scale numerical experiments are presented in Section 4.
@@
 \textbf{Theorem 2}.
-Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1$ the conditions of Theorem 1 are fulfilled.
+Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem 1 are fulfilled.
 Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1, 1\leq \nu \leq m+1,$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e.
@@
 
-	Let us show that $\bar y$ is also the limit point of the sequence of trials $\{y^k\}=\{y(x^k)\}$, generated by the IA-Dl, and the sequence $\{y^k\}$ has no new limit points.
+	Let us show that $\bar y$ is also the limit point of the sequence of trials $\{y^k\}=\{y(x^k)\}$, generated by the IA-DL, and the sequence $\{y^k\}$ has no new limit points.
 Let us carry out the proof by contradiction.
@@
 
-	At every iteration of the IA-DL method there will exist such interval $[x_{j-1}, x_j], j=j(k)$ that $\bar x \in [x_{j-1}, x_j]$.
-According to our assumption that $\bar x$ is not the limit point, there will exist such index $k$, starting from which the trial points will not fall into the interval $[x_{j-1}, x_j]$.
+	At every iteration of the IA-DL method there will exist such interval $[x_{j-1}, x_j]$, $j=j(k)$, that $\bar x \in [x_{j-1}, x_j]$.
+According to our assumption that $\bar y = y(\bar x)$ is not the limit point, there will exist such number $k$, starting from which the trial points will not fall into the interval $[x_{j-1}, x_j]$.
 That is, to conduct a new trial at the point $x^{k+1} \in [x_{t-1}, x_t]$ at every iteration such interval $[x_{t-1}, x_t], t=t(k)$ will be selected that $t \neq j$.
 
-	Therefore, in the formula (\ref{R_max}) for computing the characteristic, the second component takes the highest values, i.e. 
-$$
-	R_{glob}(t) < \rho R_{loc}(t)
-$$
+	Therefore, in the formula (\ref{R_max}) for computing the characteristic, the second component takes the highest values, i.e. $ R_{glob}(t) < \rho R_{loc}(t)$.
 Then, there will exist a limit point $x' \neq \bar x$, at that either $\nu(x') \leq m$ (i.e.
@@
 $$
-	z_j=g_{\nu}\left( y(x_j) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{1/N}, \; \nu=\nu(x_j)
+	z_j=g_{\nu}\left( y(x_j) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{1/N}, \; \nu=\nu(x_j),
 $$
 $$
-	z_{j-1}=g_{\nu}\left( y(x_{j-1}) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \; \nu=\nu(x_{j-1})
+	z_{j-1}=g_{\nu}\left( y(x_{j-1}) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \; \nu=\nu(x_{j-1}).
 $$
 
-	Since the limit point $\bar u = y(\bar x)$ belongs to the set $Y_{\epsilon}$, then
+	Since the limit point $\bar y = y(\bar x)$ belongs to the set $Y_{\epsilon}$, then
 $$
-	g_{\nu}\left( y(\bar x) \right) \leq -\epsilon_{\nu}, \; 1\leq\nu\leq m.
+	g_{\nu}\left( y(\bar x) \right) \leq -\epsilon_{\nu}, \; 1\leq\nu\leq m,
 $$
 $$
-	g_{m+1}\left( y(\bar x) \right) \leq z_{m+1}^*
+	g_{m+1}\left( y(\bar x) \right) \leq z_{m+1}^\ast.
 $$
@@
 	&=& \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	&=& \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
+%	&=& \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
 	&\geq& \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	&=& \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \geq 0
+	&=& \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \geq 0.
 	\nonumber
@@
 	&=& 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	&\geq& 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0
+	&\geq& 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
 	\nonumber
@@
 In order to construct the problems, the functions generated by the GKLS generator \cite{Gaviano2003} were used.
-This generator of multiextremal functions is often used for investigating global optimization algorithms \cite{Barkalov2018, Paulavicius2014, Sergeyev2015}.
-According to \cite{Gaviano2003}, we used simpler and more complex functions, which are  further denoted as “Simple” and “Hard”, respectively.
+This generator of multiextremal functions is often used for investigating global optimization algorithms \cite{Paulavicius2014, Sergeyev2015,Barkalov2018}.
+According to \cite{Gaviano2003}, we used simpler and more complex functions, which are  further denoted as ``Simple'' and ``Hard'', respectively.
 The complexity of the functions is determined, in particular, by the radius of the global minimizer's attraction region and by the distance between global and local minimizers.
@@
   \item the volume of the feasible subdomain $Q_{m+1}$ relative to the whole search domain $D$ is defined;
-	\item The global minimizer is known in advance and can be located at the boundary of the feasible domain;
-	\item The number of constraints active in the global minimizer can be predefined.
+	\item the global minimizer is known in advance and can be located at the boundary of the feasible domain;
+	\item the number of constraints active in the global minimizer can be predefined.
 \end{itemize}
@@
 	In the experiments that we have conducted, the number of constraints in the problems was equal to 2, and one of these constraints was active at the point of the global minimum.
-For example, the contour plots of the objective functions and constraints for $N=2$ are presented in Figs.
+For example, the level lines of the objective functions and constraints for $N=2$ are presented in Figs.
 \ref{fig:examples_simple} and \ref{fig:examples_hard}; the feasible domains are highlighted by green.
 It can be seen that the feasible domains in these problems are non-convex and non-connected, and the global minimizers are located at the boundaries of the feasible domains.
-Also, one can see that the problems of the "Hard" class have, in general, a more complex structure: the global minimizers are located at the boundaries of the feasible domains of smaller areas whereas in the problems of the “Simple” class the optimizers are located at the boundaries of the subdomains of greater areas.
+Also, one can see that the problems of the ``Hard'' class have, in general, a more complex structure: the global minimizers are located at the boundaries of the feasible domains of smaller areas whereas in the problems of the ``Simple'' class the optimizers are located at the boundaries of the subdomains of greater areas.
 
@@
 Each series consisted of 100 problems and was solved by the IA and IA-DL methods.
-The global minimizer $y^{\ast}$ was considered to be found, if the algorithm generated a trial point $y^k$ in the vicinity of the global minimum i.
-e.
-$\| y^{\ast} - y^k \| < \varepsilon \|b - a\|$ where $a$ and $b$ are the boundaries of the search domain $D$.
+The global minimizer $y^{\ast}$ was considered to be found, if the algorithm generated a trial point $y^k$ in the vicinity of the global minimum, i.e.
+$\| y^{\ast} - y^k \| < \varepsilon \|b - a\|$, where $a$ and $b$ are the boundaries of the search domain $D$.
 To reduce the dimensionality, the evolvent constructed with the parameter $m=10$ was used in both algorithms.
 The relative accuracy of the search for the solution at $N = 2, 3$ was $\epsilon=0.01$ and for the solution at $N=4,5$  it was $\epsilon=0.03$.
-The reservation parameters $\epsilon_{\nu}=\mu_{\nu} \delta,1 \leq \nu \leq 2$ were set according to formula (\ref{epsilon_nu}) with $\delta=0.01$.
+The reservation parameters $\epsilon_{\nu}=\mu_{\nu} \delta,\; 1 \leq \nu \leq 2$, were set according to formula (\ref{epsilon_nu}) with $\delta=0.01$.
 The maximum number of iterations at $N \leq 4$ was set as $K_{max} = 10^6$, and at $N=5$ as $K_{max}=7 \cdot 10^6$.
@@
 	Since all functions involved in the statement of each particular problem belong to the same class, equal parameters $r_{\nu}=4.1$ were set for the problems of the “Simple” class and $r_{\nu}=6.2$ for the problems of the “Hard” class.
-When solving the problems of these classes by the IA-DL method, the values of the parameters $r_{\nu}$ given above were selected for the upper estimate of the Lipschitz constants i.
-e.
+When solving the problems of these classes by the IA-DL method, the values of the parameters $r_{\nu}$ given above were selected for the upper estimate of the Lipschitz constants, i.e.
 the values $r_{\nu}^{glob}=4.1$ and $r_{\nu}^{glob}=6.2$ were set, which were supplemented by various values of $r_{\nu}^{loc}$.
@@
 		\end{tabular}
-	\caption{Average number of iterations performed by the IA and IA-DL methods on “Simple” class}
+	\caption{Average number of iterations performed by the methods on “Simple” class}
 	\label{tab:average_number_of_iterations_simple}
@@
 		\end{tabular}
-	\caption{Average number of iterations performed by the IA and IA-DL methods on “Hard” class}
+	\caption{Average number of iterations performed by the methods on “Hard” class}
 	\label{tab:average_number_of_iterations_hard}
@@
 In applied global optimization problems, the functions are usually defined as the “black boxes”.
-Consequently, the values of the Lipschitz constant for them are unknown \textit{a priori}, which makes their estimation one of the key problems when constructing the Lipschitz optimization methods.
+Consequently, the values of the Lipschitz constant for them are unknown a priori, which makes their estimation one of the key problems when constructing the Lipschitz optimization methods.
 

### 2020-01-27 2019/NUMTA_Barkalov/Special Issure/article.tex
figures are corrected

@@
 
-\begin{figure}[h]
-	\center{\includegraphics[width=0.7\linewidth]{epsilon_reserved_solution}}
+\begin{figure}%[h]
+	\center{\includegraphics[width=0.6\linewidth]{epsilon_reserved_solution}}
 	\caption{$\varepsilon$-reserved solution.}
@@
 \ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
-\begin{figure}[h]
-	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=4} \\ (a)}
+\begin{figure}%[h]
+	\begin{minipage}[h]{0.45\linewidth}
+		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=4} \\ {\footnotesize(a)}}
 	\end{minipage}
 	\hfill
-	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=5} \\ (b)}
+	\begin{minipage}[h]{0.45\linewidth}
+		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=5} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{stronginC3_IA} \\ a)}
+		\center{\includegraphics[width=\linewidth]{stronginC3_IA} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{stronginC3_IA-DL} \\ b)}
+		\center{\includegraphics[width=\linewidth]{stronginC3_IA-DL} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Simple_59_IA} \\ a)}
+		\center{\includegraphics[width=\linewidth]{Simple_59_IA} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Simple_59_IA-DL} \\ b)}
+		\center{\includegraphics[width=\linewidth]{Simple_59_IA-DL} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Hard_7_IA} \\ a)}
+		\center{\includegraphics[width=\linewidth]{Hard_7_IA} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Hard_7_IA-DL} \\ b)}
+		\center{\includegraphics[width=\linewidth]{Hard_7_IA-DL} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 To solve the problems shown in Figs.
-\ref{fig:examples_simple} and \ref{fig:examples_hard}, the IA method required 643 and 1166 iterations whereas the IA-DL method solved similar problems for 373 and 712 iterations, respectively.
+\ref{fig:examples_simple} and \ref{fig:examples_hard}, the IA method required 643 and 1132 iterations whereas the IA-DL method solved similar problems for 373 and 443 iterations, respectively.
 
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_3} \\ a)}
+		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_3} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_3} \\ b)}
+		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_3} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_5} \\ a)}
+		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_5} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_5} \\ b)}
+		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_5} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Decision_rule_gklssC_4} \\ a)}
+		\center{\includegraphics[width=\linewidth]{Decision_rule_gklssC_4} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Decision_rule_gklshC_4} \\ b)}
+		\center{\includegraphics[width=\linewidth]{Decision_rule_gklshC_4} \\ {\footnotesize(b)}}
 	\end{minipage}

### 2020-01-27 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
Статья по MCO по результатам конференции WCGO

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{academicons}
%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Optimization Letters}
%
\begin{document}

\title{A Computationally Efficient Approach for Solving Lexicographic Multicriteria Optimization Problems\thanks{This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance.''}
}

\titlerunning{Efficient Approach for Solving Multicriteria Optimization Problems}        % if too long for running head

\author{Victor Gergel %\href{https://orcid.org/0000-0002-4013-2329}{\includegraphics{ORCID.png}}
%\orcidID{0000-0002-4013-2329} 
\and
Evgeniy Kozinov
%\orcidID{0000-0001-6776-0096} 
\and
Konstantin Barkalov
%\orcidID{0000-0001-5273-2471}
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{               
V. Gergel \at
Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
\email{gergel@unn.ru} 
\and
E. Kozinov \at
Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
\email{evgeny.kozinov@itmm.unn.ru} 
\and
K. Barkalov \at
Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
\email{barkalov@vmk.unn.ru}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
This paper proposes a computationally efficient approach for solving complex lexicographic multicriteria optimization problems in which efficiency criteria can be multiextremal and computing the values of criteria can be time-consuming. It is also assumed that in the course of calculations, how problems are formulated may change and consequently, solving dynamically defined sets of multicriteria optimization problems may become necessary. The proposed approach is based on reducing multidimensional problems to one-dimensional global optimization problems, utilizing efficient global search algorithms, and reusing the search information obtained in the process of calculations. Results of computational experiments demonstrate that the proposed approach can significantly reduce the computational complexity of solving multicriteria optimization problems.
\keywords{Multicriteria optimization \and Lexicographic ordering \and Global optimization \and Dimensional reduction \and Search information \and Computational complexity}
\end{abstract}

\section{Introduction}
\label{sec:1}

Multicriteria optimization (MCO) problems are among the most common statements used to address decision-making problems and are widely found in applications. Multicriteria optimization is an area of vigorous scientific research, and during this time a large number of efficient methods of multicriteria optimization have been proposed and many practical problems have been solved (see, e.g., monographs \cite{c1,c2,c3,c4}) and reviews of scientific and practical results in this field \cite{c6,c7,c8,c9}.

A key feature of MCO problems is the absence, in most cases, of a single decision which would best fulfill all efficiency criteria, due to possible inconsistencies. As a result, solving MCO problems usually requires finding several compromise (efficient and non-dominated) decisions that cannot be improved without damaging the effectiveness of some efficiency criteria in the MCO problem.

Among the approaches which have been developed for solving MCO problems, lexicographic optimization methods can be singled out, when some ordering of criteria by importance is implemented and optimization is carried out based on their ordering \cite{c3}. Iterative methods represent another approach \cite{c6,c10}, when the decision maker is actively involved in the process of selecting decisions. An actively appropriated approach consists in developing and applying evolutionary algorithms based on the simulation of certain natural phenomena to solve MCO problems \cite{c10,c11,c12,c13}. Among the widely used endeavors in solving MCO problems is the scalarization approach, which applies methods involving the convolution of criteria to a single scalar criterion (see, e.g. \cite{c2,c14}).

This work is devoted to solving lexicographic multicriteria optimization (MCOlex) problems that arise when designing complex technical objects and systems. In such applications, efficiency criteria may have a complex multiextremal form, and calculating criteria values may require large amounts of computations. In such conditions, finding even one efficient decision requires significant calculations, while determining several (or a whole set) of efficient decisions becomes an issue of great computational complexity.

The properties listed above highlight a key feature of the MCOlex problems   high computational complexity. One promising attempt to solve such problems is using the model-based approach, when, after a small number of calculations determining the values of time-consuming criteria and, fast-computable approximating functions are constructed \cite{c15,c16}. This approach is quite efficient; however, it is difficult to construct good approximations given the multiextremal behavior of efficiency criteria.

This paper proposes a solution to the time-consuming class of MCOlex problems by using an approach based on the following main points. First of all, the problem is reduced to solving a sequence of global optimization problems with nonlinear constraints (GCO) \cite{c2,c14}. Then, efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization are applied for solving the GCO problems \cite{c17,c18}. And finally, the search information obtained in the process of solving the MCOlex problem is fully utilized. In general, the developed approach allows us to significantly reduce the amount of executed computations, up to just a few iterations when searching for new efficient decisions.

The further structure of the paper is as follows. Section \ref{sec:2} presents a new class of optimization problems -- multistage lexicographic multicriteria optimization (MMCOlex) problems -- whose solution is reduced to solving a sequence of global optimization problems with nonlinear constraints. Section \ref{sec:3} presents the basics of the approach that has been developed: reducing multidimensional MCOlex problems to one-dimensional global optimization problems, applying efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization, and reusing the search information obtained in the process of calculations. Section \ref{sec:4} contains the results of numerical experiments. The paper concludes by discussing the results that were obtained and suggests some main directions for further research.

\section{Problem statement}
\label{sec:2}

The multicriteria (or vector) optimization (MCO) problem can be defined as follows:
\begin{equation}\label{eq:1}
f(y) = (f_1(y), f_2(y), \dots , f_s(y)) \to min,  y \in D,
\end{equation}
where $f(y) = (f_1(y), f_2(y), \dots , f_s(y))$ is the vector efficiency criterion, $y = (y_1, y_2, \dots , y_N)$ is the vector of variable parameters, and $N$ is the dimension of the multicriteria optimization problem to be solved. The search domain $D$ defines the set of possible parameter values and is usually an $N$-dimensional hypercube
\begin{equation}\label{eq:2}
D  = \{ y\in R^N: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}
\end{equation}
for the given boundary vectors $a$ and $b$.

Without a loss of generality, it is assumed that the values of the efficiency criteria are not negative and their decrease corresponds to an increase in the effectiveness of the considered decisions $y \in D$. Problem (\ref{eq:1}) is considered in relation to the most difficult decision-making problems in which the efficiency criteria $f_i(y)$, $1 \leq i \leq s$, can be significantly multiextremal, and the procedures for calculating the values of efficiency criteria at points of the search domain $y \in D$ can be computationally expensive. It is also assumed that the criteria $f_i(y)$, $1 \leq i \leq s$, satisfy the Lipschitz condition
\begin{equation}\label{eq:2l}
|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, y',y''\in D, 1 \leq i \leq s
\end{equation}
where $L_i$ is the Lipschitz constants for the function $f_i(y)$, $1 \leq i \leq s$, and $\|*\|$ denotes the Euclidean norm in $R^N$. Fulfillment of the Lipschitz condition means that for small variations of the decision $y \in D$, corresponding changes in the values of the criteria $f_i(y)$, $1 \leq i \leq s$, are limited.

In problem (\ref{eq:1}), the efficiency criteria are usually contradictory and there is no decision $y \in D$, that would provide the best (smallest) values for all criteria at the same time, i.e.
\begin{equation}\label{eq:3}
\nexists y^*\in D: f_i(y^*) = \min_{y \in D} {f_i (y)} , 1 \leq i \leq s.
\end{equation}

If the relation (\ref{eq:3}) is valid when solving the MCO problem, then decisions $\widetilde{y}^* \in D$ can be determined, at which the values of the criteria cannot be improved without deterioration of the values of some criteria $f_i(y)$, $1 \leq i \leq s$, that is
\begin{equation}\label{eq:4}
\nexists y'\in D: (f_i(y') \leq f_i(\widetilde{y}^*)), 1 \leq i \leq s, (\exists j : f_j(y') < f_j(\widetilde{y}^*)).
\end{equation}

Such non-improvable decisions are called Pareto \textit{efficient} or \textit{optimal}.

The Pareto set $P(f,D)$ of all efficient decisions can be quite large, which makes it difficult for the decision maker to analyze. A possible way to reduce the number of efficient decisions under consideration is to assume that the effectiveness criteria are ordered in terms of importance, which is often the case in practical applications.

Without a loss of generality, we will further assume that the criteria $f_i(y)$, $1 \leq i \leq s$, are ordered in importance according to their enumeration in (\ref{eq:1}). The ordering of the criteria determines the relative linear order in the search domain $D$, that is
\begin{equation}\label{eq:4d}
y' \prec y'' \Leftrightarrow \exists i, 1 \leq i \leq s: (f_i(y')<f_i(y'')) \wedge [ \forall j, 1 \leq j <i \Rightarrow (f_j(y')=f_j(y'')) ].
\end{equation}

As a result, problem (\ref{eq:1}) is reduced to the lexicographic multicriteria optimization problem (MCOlex)
\begin{equation}\label{eq:5}
f(y) = (f_1(y), f_2(y), \dots , f_s(y)) \to min_{lex},  y \in D,
\end{equation}
the solution of which is carried out in stages: initially, the first (most important) criterion is optimized; next, if the solution is not the only one, then the second criterion is optimized on the set of solutions of the first stage, and so on.
\begin{equation}\label{eq:6}
D_1=Arg \min_{y \in D}{f_1(y)},\dots, D_i=Arg \min_{y \in D_{i-1}}{f_i(y)},1 \leq i \leq s,D_0=D,
\end{equation}
where \textit{Arg} means the set of all decisions that achieve the minimum value of the optimized criterion. It should be noted that the sequence of steps in (\ref{eq:6}) may not be fully completed if at some stage $l$, $1 \leq l \leq s$, the set $D_l$ degenerates and contains only a single decision $y \in D$.

The set of efficient decisions $P_{lex}(f,D)$ obtained as a result of (\ref{eq:6}) is a subset of the Pareto domain $P(f,D)$ and can contain from one to several decisions $\widetilde{y}^* \in D$ in the case that the minimum value of the last criterion $f_s(y)$ is achieved at several points in the search domain $D$. Such a sharp reduction in the set of the efficient decisions being considered may be undesirable. Expanding the set of $P_{lex}(f,D)$ decisions can be achieved through somewhat weakening the ``strict'' lexicographic order and by allowing, at each stage of the computational scheme (\ref{eq:6}), the choice of decisions in a certain neighborhood of the minimum values of the optimized criterion, that is
\begin{equation}\label{eq:7}
\begin{matrix}
f^*_1 =\min_{y \in D}{f_1(y)},  \\
f^*_i =\min_{y \in D}{f_i(y)},  f_j(y) \leq f^*_j + \delta_j, 1 \leq j < i, 1 < i \leq s, \\
P_{lex}^\delta(f,\delta,D) = Arg \min_{y \in D}{f_s(y)},f_j(y) \leq f^*_j + \delta_j, 1 \leq j < i, 1 < i \leq s.
\end{matrix}
\end{equation}

This approach is also widely known as the method of successive concessions (MSC) \cite{c3,c4,c5}. The choice of feasible concessions $\delta_i$, $1 \leq i < s$ in the computational scheme (\ref{eq:7}) allows one to obtain any efficient decision $\widetilde{y}^* \in P_{lex}(f,D)$ and take into account the peculiarities of the MCOlex problem being solved. At the same time, this approach makes it necessary, at each stage of scheme (\ref{eq:7}), to solve more complex global optimization problems with nonlinear constraints.

It should also be noted that during the calculation process, it may be necessary to change the selected values of the feasible concessions $\delta_i$, $1 \leq i < s$; the concessions may be quite rigid or, conversely, excessively large. In general, one may need to change the order of importance in efficiency criteria. Such assumptions necessitate a more general formulation of how to solve the MCOlex problem and to provide possible solutions to many problems of the form (\ref{eq:8})
\begin{equation}\label{eq:8}
\mathbb{P}_t=\{ P_1,P_2,\dots,P_t\},
\end{equation}
which can be changed dynamically during calculations by adding new or deleting existing multicriteria optimization problems. In general, such an approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex).

\section{The Approach: Key Ideas and Methods}
\label{sec:3}

The extended formulation of lexicographic multicriteria optimization problems (\ref{eq:7})-(\ref{eq:8}) involves multiple solutions of multiextremal optimization problems with nonlinear constraints. Problems of this type are computationally time-consuming and subject to the ``curse of dimensionality'' -- computational complexity increases exponentially with increasing dimensionality. The proposed approach for solving such problems is based on three key ideas:
\begin{itemize}
  \item Reduction of multidimensional problems of multiextremal optimization to problems of one-dimensional global search by using mappings based on Peano curves;
  \item Using the index method of global optimization, which allows solving optimization problems with constraints without using penalty functions;
  \item Repeated use of search information obtained in the process of computing, for solving multiple global optimization problems.
\end{itemize}

\subsection{Dimension reduction of multidimensional multicriteria optimization problems}

The search for numerical estimates of globally optimal points involves the construction of grids covering the search domain $D$ (see, e.g., \cite{c17,c18,c19,c20,c21,c22,c23,c24,c25}). This key feature of global optimization problems leads to the ``curse of dimensionality'', where the computational complexity of constructing coverage increases exponentially as the dimensions of optimization problems are enlarged. Significantly reduced complexity can be achieved by using heterogeneous grids that are denser only in the vicinity of globally optimal points. Such heterogeneous coverage of the search domain $D$ can be constructed adaptively in the following manner: when choosing points for the next iterations, the global search is performed taking into account all the available search information (points of previous search iterations and values of the optimized function at these points), that is
\begin{equation}\label{eq:9}
y^{k+1}=G_k(\omega_k), \omega_k=\{(y^i,f^i=f(y^i)): 1 \leq i \leq k \},
\end{equation}
where $k$ is the number of the global search iterations being performed, $G_k$ is the decisive rule of the applied global search algorithm, according to which the point $y^{k+1}$ of the next iteration is selected, and $\omega_k$ is the search information available at the $k$ step of the global search.

Adaptive selection of points $y^{k+1} \in D$ at the iterations of the global search involves the analysis of volumetric multidimensional search information $\omega_k (k \gg 1)$, which determines the high computational complexity of the decisive rules $G_k$,$k=1,2,\dots$ of global search algorithms. This computational complexity can be significantly simplified by reducing the optimization problems to be solved using Peano \textit{space-filling curves} or \textit{evolvents} $y(x)$, that uniquely and continuously map the interval $[0,1]$ to an $N$-dimensional hypercube D (see, e.g., \cite{c17,c18,c23}). As a result of this reduction, the initial multidimensional multicriteria optimization problem (\ref{eq:5}) is reduced to a one-dimensional problem:
\begin{equation}\label{eq:10}
f(y(x)) = (f_1(y(x)), f_2(y(x)), \dots , f_s(y)(x)) \to min_{leq},  x \in [0,1].
\end{equation}

It should be noted that the one-dimensional functions obtained as a result of the reduction satisfy the uniform H\"older condition (see \cite{c17,c18}), that is
\begin{equation}\label{eq:11}
|f_i (y(x'))-f_i (y(x';))| \leq H_i \|x'-x''\|, x',x''\in [0,1], 1 \leq i \leq s,
\end{equation}
where the constants $H_i$ are determined by the ratio $H_i=2L_i\sqrt{N+3}$, $1 \leq i \leq m$, $L_i$ are the Lipschitz constants from (\ref{eq:2l}), and $N$ is the dimension of the optimization problem (\ref{eq:1}).

Using dimension reduction the search information $\omega_k$ can be converted from (\ref{eq:9}) to the form
\begin{equation}\label{eq:12}
A_k=\{(x_i, y(x_i), f_i=f(y(x_i)): 1 \leq i \leq k \},
\end{equation}
in which, in addition to the reduced representation, the data is arranged in ascending order\footnote{Data ordering is reflected by using a subscript.} of points $x_i$, $1 \leq i \leq k$, that is
\begin{equation}
x_1 < x_2 < \dots < x_k	
\end{equation}
for a more efficient execution of global search algorithms.

Dimension reduction using Peano curves is widely used in the development of efficient global search algorithms in the framework of the information-statistical theory of multiextremal optimization \cite{c17,c18,c23,c26,c27}. Numerical methods for constructing approximations of Peano curves (\textit{evolvents}) with a given accuracy are considered in \cite{c17,c18}.

\subsection{An efficient method for solving global optimization problems with 
nonlinear constraints}

The problems solved in the framework of the computational scheme (\ref{eq:7}) are one-dimensional global optimization problems with nonlinear constraints
\begin{equation}\label{eq:13}
\min{\{\varphi(x):g_i(x)\leq 0, 1 \leq i \leq m, x\in \Delta=[0,1]\}}
\end{equation}
(the notation $g_{m+1}(x) = \varphi(x)$ will also be used later).

The presence of nonlinear constraints significantly complicates the complexity of solving global optimization problems; the resulting solutions must belont to a feasible domain
\begin{equation}\label{eq:14}
\Delta_g  = \{ x\in \Delta=[0,1]:g_i(x)\leq 0,   1 \leq i \leq m \}.
\end{equation}

To reduce the complexity of solving optimization problems with constraints, simpler cases are often highlighted; for example, problems with linear or quadratic constraints are selected. Various methods of approximating complex constraints with the aid of simpler forms (linear, convex, etc.) are widely used. However, the penalty function method is most commonly used, when calculating in an unfeasble region $\Delta \setminus \Delta_g$, some ``penalty'' are added to the values of the minimized function. But the penalty function method often involves repeatedly solving the optimization problem to determine a sufficient value of the penalty coefficient. More complete information on methods for solving optimization problems with constraints can be obtained, for example, in monographs \cite{c18,c26,c27}.

The proposed approach is based on the original method of separate accounting for constraints developed in the framework of the information-statistical theory of multiextremal optimization \cite{c18}. The key idea of the approach consists in constructing a certain integral objective function without constraints, the solution of which also provides a solution to the initial constrained optimization problem (\ref{eq:13}); a more detailed presentation of this method is given below.

We apply a sequential scheme to obtain the constraint values $g_i(y(x)) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$, according to which the constraint values are calculated strictly in the order of their numbers, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme). The number of constraints in which the values were calculated will be referred below as the index $\nu=\nu(x)$, that is
\begin{equation}\label{eq:15}
\nu(x) : g_\nu(x)>0, g_j(x)\leq 0, 1 \leq j\leq \nu - 1, 1\leq \nu = \nu(x) \leq m
\end{equation}
 (when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index concept allows us to determine the integrated function for the problem (\ref{eq:13}) (see Fig. \ref{fig:1})
\begin{equation}\label{eq:16}
F(x)=g_\nu(x), \nu=\nu(x),
\end{equation}
defined and computable everywhere in $[0,1]$. Its value at the point $x\in [0,1]$ is either the value of the left side of the constraint violated at this point (the case when $\nu \leq m$), or the value of the minimized function (the case when $\nu = m + 1$). The process of calculating the value of $F(x)$, $x\in [0,1]$, is completed either as a result of establishing the inequality $g_j(x)>0$, or as a result of reaching the value $\nu(x)=m+1$ (below, this procedure is called a \textit{trial}).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig1}
  \caption{An integrated function for the problem (\ref{eq:13}) using a sequential constraint calculation scheme: a) the function $F(x)$ from (\ref{eq:16}), b) the function $\Phi(x)$ from (\ref{eq:17})}
  \label{fig:1}
\end{figure}

The function $F(x)$ consists of separate fragments of the constraints $g_i(x) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$. For the homogeneity of the components, the function $F(x)$ is then converted to the form 
\begin{equation}\label{eq:17}
\Phi (x) = 
 \begin{cases}
   g_\nu(x) / H_\nu, & \nu < M, \\
   (g_M(x)-z^*_M) / H_M, & \nu = M,
 \end{cases}
\end{equation}
where $M$ is the constraint number for which there are no feasible points (the maximum value of the index is $\nu(x)$), $z^*_M$ is the value of the minimum constraint violation $g_M(x)$\footnote{If $M=m+1$, then $z_M^*$ is the minimum value of the function $\varphi(x)$.}, and $H_\nu$, $1\leq \nu \leq M$, are the H\"older constants from (\ref{eq:11}). As a rule, these values are a priori unknown, but when performing calculations, instead of these values, their adaptive estimates obtained in the process of calculations can be applied. For this, the search information $A_k$ from (\ref{eq:12}) should be expanded with data
\begin{equation}\label{eq:18}
A_k=\{(x_i, z_i, \nu_i, y(x_i), f_i=f( y(x_i)): 1 \leq i \leq k \},
\end{equation}
where $z_i= \Phi(x_i)$, $1 \leq i \leq k$. Then estimates of the required values can be determined as follows
\begin{equation}\label{eq:19}
M^k= \max{\{\nu=\nu(x_i), 1\leq i \leq k \}},
\end{equation}

\begin{equation}\label{eq:20}
z^k_\nu = 
 \begin{cases}
   0, & \nu < M, \\
   \min{\{g_\nu(x_i):\nu=\nu(x_i), 1\leq i \leq k \}}, & \nu = M,
 \end{cases}
\end{equation}

\begin{equation}\label{eq:21}
h^k_l=\max{\{|z_i-z_j|/\sqrt[N]{x_i-x_j},\nu_i=\nu_j=l,i>j,1\leq l\leq M^k\}}
\end{equation}

If $h_l^k$ in (\ref{eq:21}) turns out to be indefinite or equal to zero, then $h_l^k=1$ is taken. For the initial values of these quantities, $M^0 = H_l^0=1$, $z_l^0=0$, $1 \leq i \leq m+1$ can be taken.

The general computational scheme of the global search algorithm for multiextremal optimization problems with nonlinear constraints \footnote{This method is also known as the index method -- see \cite{c18}.} (AGCS) is as follows \cite{c18}.

The first trial is carried out at an arbitrary internal point $x^1\in (0,1)$. The calculation results are constantly added to the set $A_k$ from (\ref{eq:12}); in addition, into $A_k$ are also added the points $x_0=0$ and $x_{k+1}=1$ (the values $z_0$ and $z_{k+1}$ are not defined and not used) for the convenience of subsequent formulations. The choice of the point $x^{k+1}$, $k\geq 1$ of any next trial is determined by the following rules.

\textit{Rule 1.} For each interval $(x_{i-1},x_i)$, $1 \leq i \leq k+1$, in the set $A_k$ calculate the characteristic $R(i)$, where
\begin{equation}\label{eq:22}
\begin{matrix}
R(i) = 
 \begin{cases}
   \rho_i + \frac{(z_i-z_{i-1})^2}{r^2_\nu h^2_\nu \rho_i} - 2 \frac{z_i+z_{i-1}-2z^*_\nu}{r_\nu h_\nu}, & \nu=\nu(x_{i-1})=\nu(x_i), \\
   2 \rho_i - 4\frac{z_i-z^*_\nu}{r_\nu h_\nu}, & \nu=\nu(x_{i-1})<\nu(x_i), \\
2 \rho_i - 4\frac{z_{i-1}-z^*_\nu}{r_\nu h_\nu}, & \nu=\nu(x_{i-1})>\nu(x_i), 
 \end{cases} \\
\rho_j = \sqrt[N]{x_i-x_{i-1}}, 1\leq j\leq k+1.
\end{matrix}
\end{equation}

The constants $r_\nu>1$, $1\leq \nu \leq m+1$, are the parameters of the algorithm. The products $r_\nu h_\nu$ are used as an estimate of the H\"older constants $H_\nu$, $1\leq \nu \leq m+1$ from (\ref{eq:11}).

\textit{Rule 2.} Determine the interval $(x_{t-1},x_t)$ that corresponds to the maximum characteristic
\begin{equation}\label{eq:23}
R(t) = \max{\{R(i):1 \leq i \leq k+1\}}.
\end{equation}

\textit{Rule 3.} Execute a new trial at the point $x^{k+1} \in (x_{t-1},x_t)$, determined in accordance with the expression
\begin{equation}\label{eq:24}
x^{k+1} = 
 \begin{cases}
   \frac{x_t+x_{t-1}}{2}, & \nu(x_{t-1}) \neq \nu(x_t), \\
   \frac{x_t+x_{t-1}}{2} + sign(z_t - z_{t-1}) \frac{1}{2r_\nu}[ \frac{|z_t - z_{t-1}|}{h_\nu}^N ], & \nu = \nu(x_{t-1}) = \nu(x_t).
 \end{cases}
\end{equation}

Iterations of the algorithm are terminated if the stop condition is satisfied
\begin{equation}\label{eq:25}
\rho_t\leq \varepsilon,
\end{equation}
where $t$ from (\ref{eq:23}) and $\varepsilon > 0$ is the given accuracy.

Results of using the AGCS algorithm to solve the test problem from Fig. \ref{fig:1} with the parameters $r_1=r_2=r_3=2$ and $\varepsilon=10^5$are shown in Fig. \ref{fig:2}.  The coordinates of the trial points performed by the algorithm in the process of solving the problem are marked in Fig. \ref{fig:1} by three rows of vertical strokes. The strokes in the upper row correspond to points with a unit index, the second row to points whose indices equal to 2; the points marked with strokes in the bottom row belong to the feasible domain. The coordinates of trials performed at close points are marked with a dark rectangle. In total, the value of the first constraint was computed 147 times, the value of the second constraint -- 84 times, and the value of the minimized function was calculated only 35 times.

\begin{figure}
  \centering
  \includegraphics[width=0.6\linewidth]{fig2}
  \caption{Results of using the AGCS algorithm to solve the test problem}
  \label{fig:2}
\end{figure}

Various modifications of this algorithm and the corresponding theory of convergence are presented in \cite{c18}. For instance, the following theorem is valid.

\begin{theorem}
If the AGCS algorithm is used to solve problem (\ref{eq:13}), and the following conditions are satisfied:
\begin{enumerate}
	\item the functions $g_j$, $1 \leq j \leq m+1$,  satisfy the H\"older condition of (\ref{eq:11}) with the constants $H_j$, $1 \leq j \leq m+1$,
	\item for the quantities $h_\nu$ from (\ref{eq:21}), starting from a certain step, the inequalities are valid
\begin{equation*}
r_\nu h_\nu>2H_\nu, 1\leq \nu \leq m+1,
\end{equation*}
\end{enumerate}
then the set of limit points of the sequence $\{x^k\}$, generated by the algorithm coincides with the set of global minimum points of the problem (\ref{eq:13}) with $\varepsilon=0$ in the stop condition (\ref{eq:25}).
\end{theorem}

\subsection{Acceleration of computations based on the re-use of search information}

As noted earlier, the MCOlex problem at each stage of the successive concessions method (\ref{eq:7}) is a global optimization problem with nonlinear constraints (\ref{eq:13}). Generally, the solution to such problems must start at the beginning, and multiple solutions of such global optimization problems may require extensive calculations. However, the presence of search information $A_k$ from (\ref{eq:18}) allows us to bring the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:7}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$,  of the initial problem (\ref{eq:1}) \cite{c30,c31}. This re-use of search information, to find the next efficient decision, reduces the amount of computations needed to solve each next optimization problem to just a few iterations; see Section \ref{sec:4} for the results of computational experiments.

Moreover, using search information $A_k$ allows to reduce the multistage solution of the MCO problem in accordance with the scheme (\ref{eq:7}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:7}), namely 
\begin{equation}\label{eq:26}
P_{\delta lex} (f,\delta,D)=Arg \min_{y\in D}{f_s(y)}, f_j(y)\leq f_j^*+\delta_j, 1 \leq j < s,
\end{equation}
where for the unknown a priori values $f_j^*$, $1 \leq j < s$, can be used the estimates $z_j^k$, $1 \leq j < s$, from (\ref{eq:7}) obtained from the search information $A_k$,  and for the constants $\delta_j$, $1 \leq j<s$, we can use the values normalized to the range of constraints values, that is
\begin{equation}\label{eq:27}
	\begin{matrix}
	\delta_j=z_j^k+\theta_j (g_j^k-z_j^k ), 0\leq \theta_j\leq 1,1\leq j < s,\\
	g_j^k= \max{\{ g_j (x_i ) : \nu(x_i)=j,1\leq i\leq k \}}.
	\end{matrix}
\end{equation}

The applicability of such a one-step scheme (\ref{eq:26})-(\ref{eq:27}) instead of the scheme (\ref{eq:7}) is computationally evaluated in Section \ref{sec:4}.

With a complete scheme for calculating criteria values, i.e. when at the iteration points $x_i$, $1 \leq i \leq k$,  the values of all criteria $f_j (x_i)$, $1 \leq j \leq s$, $1 \leq i \leq k$, are calculated, the effect of using the search information $A_k$ can be even more significant. In this case, the search information can also be used when changing the values of the concessions $\delta_j$, $1 \leq j < s$, and thus, the solution of the next problem from the $\mathbb{P}_t$ family (\ref{eq:8}) can be carried out repeatedly using the results of all previously performed calculations. Results of computational experiments (see Section \ref{sec:4}), demonstrate that the volume of computations performed, based on reusing search information, is reduced by a factor of no less than 6.

The AGCS algorithm, supplemented by the ability to use search information to solve multiple MCOlex  problems, will be referred to below as the global search algorithm for multistage solutions of a set of multiextremal optimization problems with nonlinear constraints (MAGCS).

\section{Results of computational experiments}
\label{sec:4}

The computational experiments were carried out on the Lobachevsky supercomputer at Nizhny Novgorod State University (the operating system CentOS 6.4, the management system SLURM). A supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM processors. Each CPU is 8-core; that is, there are 16 CPU cores available on the node. The Intel C++ 14.0.2 compiler was used to obtain an executable program code. The numerical experiments were performed using the Globalizer system \cite{c32}.

The global optimization algorithms used within the framework of the proposed approach have proven to be efficient in the process of conducting test computational experiments and have been widely used in solving practical optimization problems (see, e. g., \cite{c33,c34}). The results of computational experiments in solving multicriteria optimization problems are given below.

In the first series of computational experiments, the developed MAGCS algorithm was compared with several multicriteria optimization algorithms. To make a comparison, we used the test two-criteria problem proposed in \cite{c35}:
\begin{equation}\label{eq:28}
f_1 (y)=(y_1-1) y_2^2+1,f_2 (y)=y_2, 0\leq y_1,y_2 \leq 1.
\end{equation}

Solving the MCO problem was understood to mean constructing a numerical approximation of the Pareto domain. To assess the quality of the approximation, the completeness and uniformity of the Pareto domain coverage were compared, using the following two indicators \cite{c35,c36}:
\begin{itemize}
	\item 	The hypervolume index (HV). This indicator characterizes the approximation of the Pareto domain in terms of completeness; a higher value corresponds to a more complete coverage of the Pareto domain.
	\item The distribution uniformity index (DU). This indicator characterizes the uniformity of Pareto domain coverage; a lower value corresponds to a more uniform coverage of the Pareto domain.
\end{itemize}

In the framework of this experiment, five multicriteria optimization algorithms were compared: the Monte-Carlo (MC) method, the genetic algorithm SEMO from the PISA library \cite{c9,c36}, the Non-uniform coverage (NUC) method \cite{c35}, the bi-objective Lipschitz optimization (BLO) method \cite{c36}, and the MAGCS algorithm proposed in this paper. The results of solving problem (\ref{eq:28}) for all these methods (except for MAGCS) were obtained in \cite{c36}.

For MAGCS, 50 problems (\ref{eq:26}) were solved for different values $\theta_j$, $1 \leq j<s$, uniformly distributed in the interval [0,1]. The accuracy from (\ref{eq:25}) $\varepsilon=0.05$ and the reliability from (\ref{eq:22}) $r=3.0$ were used. The results of the executed experiments are presented in Table \ref{tab:1}.

\begin{table}[ht]
\centering
\caption{The numerical results of solving the multicriteria optimization problems from (\ref{eq:28})}
\label{tab:1}
\begin{tabular}{cccccc}
\hline
Method                                                                                     & MC    & SEMO  & NUC   & BLO   & \textbf{MAGCS} \\ \hline
Number of executed iterations                                                              & 500   & 500   & 515   & 498   & \textbf{273}   \\
\begin{tabular}[c]{@{}c@{}}Number of points in \\ Pareto domain approximation\end{tabular} & 67    & 104   & 29    & 68    & \textbf{80}    \\
HV index                                                                                   & 0.300 & 0.312 & 0.306 & 0.308 & \textbf{0.314} \\
DU index                                                                                   & 1.277 & 1.116 & 0.210 & 0.175 & \textbf{0.096} \\ \hline
\end{tabular}
\end{table}

The results from the executed experiments demonstrate that the MAGCS algorithm has a significant advantage in comparison with the considered multicriteria optimization methods of, even when solving relatively simple MCO problems.

In the second series of computational experiments, two-criteria two-dimensional MCO problems were solved. Multiextremal functions obtained using the GKLS generator \cite{c37} served as the criteria for the problem. During the experiments, 100 multicriteria problems of this class were solved. In each problem, Pareto-optimal decisions were searched for 50 different values of $\theta_j$, $1\leq j < s$, uniformly distributed in the interval [0,1]; (that is, 5,000 global optimization problems were solved). The results obtained were averaged over the number of MCO problems solved.

In the computational experiments, calculations are terminated when the required accuracy was achieved. When the process was terminated, the solution was evaluated for correctness. As a means of control, we compared the solution points found by MAGCS and the points calculated from the Pareto boundary, taking into account the selected values of $\theta_j$, $1 \leq j < s$. The method parameters were set as foolows: the accuracy $\varepsilon = 0.02$ and the reliability $r_1=r_2=5.6$. Results of the computational experiments are presented in Table 2.

In Table 2, the first and fifth columns indicate the average number of iterations executed by the MAGCS algorithm for solving the MCOlex problems. The second and sixth columns contain the percentage of completed problems. The third, fourth, seventh and eighth columns show the values of HV and DU indicators. The last column gives values showing the reduction in the number of global search iterations that were executed when solving the MCOlex problems, by reusing search information.

\begin{table}[ht]
\centering
\caption{Results of the series of experiments to solve two-dimensional two-criteria MCO problems}
\label{tab:2}
\resizebox{\columnwidth}{!}{
\begin{tabular}{ccccccccc}
\hline
\multicolumn{8}{c}{Search information}                                    & Reduction \\ \cline{1-8}
\multicolumn{4}{c}{not used}       & \multicolumn{4}{c}{used}             &  in \\ \cline{1-8}
Method     & Problem & Avg   & Avg   & Method     & Problem & Avg   & Avg   & iteration    \\
iter.      & solved  & HV    & DU    & iter.      & solved  & HV    & DU    & number       \\ \hline
41 710.1   & 98.0\%  & 33.43 & 0.173 & 2 407.41   & 99.3\%  & 33.34 & 0.224 & 17.3         \\ \hline
\end{tabular}
}
\end{table}

Results obtained from these experiments show that the re-use of search information can reduce the amount of calculations by 17.3 times without expending additional computational resources, and according to the average HV and DU indicators, the quality of the Pareto domain remains on average at the same level.

In the third series of computational experiments, the two-criteria four-dimensional MCO problems were solved. As in the second series of computational experiments, 100 multicriteria problems were solved. In each problem, the Pareto-optimal decisions were calculated for 50 different values of $\theta_j$, $1 \leq j <s$, uniformly distributed in the interval $[0,1]$. The criteria of the MCO problems being solved were determined using the GKLS generator \cite{c37}. The method parameters were set as foolows: the accuracy $\varepsilon = 0.025$ and the reliability $r_1=r_2=5.6$. The results of computational experiments are presented in Table 3.

\begin{table}[ht]
\centering
\caption{Results of the series of experiments to solve two-criteria four-dimensional MCO}
\label{tab:3}
\resizebox{\columnwidth}{!}{
\begin{tabular}{ccccccccc}
\hline
\multicolumn{8}{c}{Search information}                                    & Reduction \\ \cline{1-8}
\multicolumn{4}{c}{not used}       & \multicolumn{4}{c}{used}             &  in \\ \cline{1-8}
Method      & Problem & Avg   & Avg   & Method     & Problem & Avg   & Avg   & iteration    \\
iter.       & solved  & HV    & DU    & iter.      & solved  & HV    & DU    & number       \\ \hline
4 536 377.9 & 83.0\%  & 30.67 & 0.335 & 709 014.9  & 94.5\%  & 30.46 & 0.405 & 6.4         \\ \hline
\end{tabular}
}
\end{table}

Results from the executed experiments show that with increased dimensions of MCO problems to be solved and a corresponding increase in the volume of calculations, the efficiency of the MAGCS algorithm remains at a high level;  its use has efficiently reduced the number of the executed iterations by 6.4 times.

\section{Conclusion}
\label{sec:5}

This paper proposes a new approach for solving computationally expensive lexicographic multicriteria optimization problems (MCOlex), in which the efficiency criteria can be multiextremal, and calculating the criteria values may require a large amount of computation. A key feature of this class of problems is the ability, during the computation process, to alter the ordering of efficiency criteria in terms of importance, which in turn necessitates a multistage solution for MCOlex problems.

Overcoming the enormous computational complexity of addressing the formulated new class of MCOlex problems is ensured by solving a sequence of global optimization problems with nonlinear constraints using efficient information-statistical methods of global optimization by using an original index constraint accounting scheme instead of the commonly used penalty functions. A core element in this approach is the ability to use all the search information obtained in the computation process, utilizing the multistage solution of MCOlex problems. Starting to solve each new stage of the solution, this search information allows us to incorporate the previously calculated values of efficiency criteria into the values of the next scalar problem of multiextremal optimization. The search information retrieved becomes part of the current optimization state and through optimization methods, results in adaptive planning of the iterations to run the global search.

Results of computational experiments demonstrate that this approach can significantly reduce the computational complexity of multistage MCOlex problem solving.

In conclusion, it must be noted that the approach described here is promising, and requires further research. Importantly, it is necessary to continue carrying out computational experiments to solve multicriteria optimization problems with a larger number of efficiency criteria and for greater dimensions in the optimization problems to be solved. It is also necessary to evaluate the possibility of parallel computing using high-performance supercomputer systems.

% Non-BibTeX users please use
\begin{thebibliography}{}
%
% and use \bibitem to create references. Consult the Instructions
% for authors for reference list style.
%

\bibitem{c1} Miettinen K.: Nonlinear Multiobjective Optimization. Springer. (1999)
\bibitem{c2} Ehrgott, M.: Multicriteria Optimization. Springer. (2nd ed., 2010)
\bibitem{c3} Collette, Y., Siarry, P.: Multiobjective Optimization: Principles and Case Studies (Decision Engineering). Springer. (2011)
\bibitem{c4} Marler, R.T., Arora, J.S.: Multi-Objective Optimization: Concepts and Methods for Engineering. VDM Verlag. (2009)
\bibitem{c5} Pardalos, P.M., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-Convex Multi-Objective Optimization. Springer. (2017)
\bibitem{c6} Marler, R. T., Arora, J. S.: Survey of multi-objective optimization methods for engineering. Struct. Multidisciplinary Optimization 26, pp. 369--395 (2004)
\bibitem{c7} Figueira,J., Greco, S., Ehrgott, M., editors.: Multiple criteria decision analysis: State of the art surveys. Springer, New York (NY). (2005)
\bibitem{c8} Zavadskas, E. K., Turskis, Z., Kildien\.e, S.: State of art surveys of overviews on MCDM/MADM methods. Technological and Economic Development of Economy, 20, pp. 165--179 (2014)
\bibitem{c9} Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications. Surv. Math. Ind. 11, pp. 1--42 (2005)
\bibitem{c10} Branke, J., Deb, K., Miettinen, K., Slowinski, R., editors.: Multi-Objective Optimization—Interactive and Evolutionary Approaches. Springer, Berlin. (2008)
\bibitem{c11} Deb, K.: Multi-Objective Optimization using Evolutionary Algorithms. Wiley, Chichester. (2001)
\bibitem{c12} Yang, X.-S.: Nature-inspired metaheuristic algorithms. Luniver Press, Frome. (2008) 
\bibitem{c13} Tan, K.C., Khor, E.F., Lee, T.H.: Multi-objective Evolutionary Algorithms and Applications. Springer-Verlag, London. (2005)
\bibitem{c14} Eichfelder, G.: Scalarizations for adaptively solving multi-objective optimization problems. Comput. Optim. Appl. 44, pp. 249--273 (2009)
\bibitem{c15} Jones, D.R.: A taxonomy of global optimization methods based on response surfaces. J. Glob. Optim. 21, pp. 345--383 (2001)
\bibitem{c16} Voutchkov, I., Keane, A.: Multi-objective optimization using surrogates. Comput. Intell. Optim. Adapt. Learn. Optim. 7, pp. 155--175 (2010)
\bibitem{c17} Strongin, R.G.: Numerical methods in multiextremal problems: information-statistical algorithms. Nauka, Moscow. (1978, in Russian) 
\bibitem{c18} Strongin, R., Sergeyev, Ya.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht. (2nd ed. 2013, 3rd ed. 2014).
\bibitem{c19} T\"orn, A., {\v Z}ilinskas, A.: Global Optimization. Lecture Notes in Computer Science, 350. Springer-Verlag, Berlin. (1989)
\bibitem{c20} Horst, R., Tuy, H.: Global Optimization: Deterministic Approaches. Springer-Verlag, Berlin. (1990)
\bibitem{c21} Zhigljavsky, A.A.: Theory of Global Random Search. Kluwer Academic Publishers, Dordrecht. (1991)
\bibitem{c22} Pint\'er, J.D.: Global optimization in Action (continuous and Lipschitz optimization: algorithms, implementations and applications). Kluwer Academic Publishers, Dortrecht. (1996)
\bibitem{c23} Sergeyev Y.D., Strongin R.G., Lera D.: Introduction to global optimization exploiting space-filling curves. Springer. (2013)
\bibitem{c24} Locatelli, M., Schoen, F.: Global Optimization: Theory, Algorithms, and Applications. SIAM. (2013)
\bibitem{c25} Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. Princeton University Press. (2016)
\bibitem{c26} Lera D., Sergeyev Ya.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and Holder constants, Communications in Nonlinear Science and Numerical Simulation, 23. pp. 328--342 (2015)
\bibitem{c27} Gergel, V.: A Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems. 2nd International Conference on Mathematics and Computers in Sciences and in Industry. pp. 13--18 (2015). DOI: 10.1109/MCSI.2015.18
\bibitem{c28} Arora, R.K.: Optimization: Algorithms and Applications. CRC Press. (2015)
\bibitem{c29} Bazaraa, M.S., Sherali, H.D., Shetty, C. M.: Nonlinear Programming: Theory and Algorithms, John Wiley and Sons. (2006)
\bibitem{c30} Gergel, V.P., Kozinov, E.A.: Accelerating multicriterial optimization by the intensive exploitation of accumulated search data. In: AIP Conference Proceedings, 1776. pp. 090003 (2016). DOI: 10.1063/1.4965367
\bibitem{c31} Gergel, V.P., Kozinov, E.A.: Efficient multicriterial optimization based on intensive reuse of search information.  J. Glob. Optim., 71(1). pp. 73--90 (2018). DOI: 10.1007/s10898-018-0624-3
\bibitem{c32} Sysoyev, A., Barkalov, K., Gergel, V.: Globalizer: A novel supercomputer software system for solving time-consuming global optimization problems. Numerical Algebra, Control and Optimization 8(1), pp. 47--62 (2018)
\bibitem{c33} Kvasov D.E., Sergeyev Y.D.: Deterministic approaches for solving practical black-box global optimization problems. Advances in Engineering Software, 80, pp.58--66 (2015)
\bibitem{c34} Modorskii, V., Gaynutdinova, D., Gergel, V., Barkalov, K.: Optimization in design of scientfic products for purposes of cavitation problems. AIP Conference Proceedings, 1738. (2016)
\bibitem{c35} Evtushenko, Yu.G., Posypkin, M.A.: A deterministic algorithm for global multi-objective optimization. Optimization Methods \& Software, 29(5), pp. 1005--1019 (2014)
\bibitem{c36} {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Adaptation of a one-step worst-case optimal univariate algorithm of bi-objective Lipschitz optimization to multidimensional problems. Commun Nonlinear Sci Numer Simulat, 21, pp. 89--98 (2015)
\bibitem{c37} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Transactions on Mathematical Software 29(4). pp. 469--480 (2003)

\end{thebibliography}

\end{document}
% end of file template.tex

### 2020-01-27 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
исправил ссылки

@@
 In general, one may need to change the order of importance in efficiency criteria.
-Such assumptions necessitate a more general formulation of how to solve the MCOlex problem and to provide possible solutions to many problems of the form (\ref{eq:8})
+Such assumptions necessitate a more general formulation of how to solve the MCOlex problem and to provide possible solutions to many problems of the form (\ref{eq:7})
 \begin{equation}\label{eq:8}
@@
 \end{equation}
-where for the unknown a priori values $f_j^*$, $1 \leq j < s$, can be used the estimates $z_j^k$, $1 \leq j < s$, from (\ref{eq:7}) obtained from the search information $A_k$,  and for the constants $\delta_j$, $1 \leq j<s$, we can use the values normalized to the range of constraints values, that is
+where for the unknown a priori values $f_j^*$, $1 \leq j < s$, can be used the estimates $z_j^k$, $1 \leq j < s$, from (\ref{eq:20}) obtained from the search information $A_k$,  and for the constants $\delta_j$, $1 \leq j<s$, we can use the values normalized to the range of constraints values, that is
 \begin{equation}\label{eq:27}

### 2020-01-27 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
a lot of minor corrections

@@
 \usepackage{amsmath}
+\usepackage{hyperref}
 %\usepackage{academicons}
@@
 
-\title{A Computationally Efficient Approach for Solving Lexicographic Multicriteria Optimization Problems\thanks{This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance.''}
+\title{A Computationally Efficient Approach for Solving Lexicographic Multicriteria Optimization Problems\thanks{This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''.}
 }
@@
 Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
-\email{barkalov@vmk.unn.ru}
+\email{konstantin.barkalov@itmm.unn.ru}
 }
@@
 Results of computational experiments demonstrate that the proposed approach can significantly reduce the computational complexity of solving multicriteria optimization problems.
-\keywords{Multicriteria optimization \and Lexicographic ordering \and Global optimization \and Dimensional reduction \and Search information \and Computational complexity}
+\keywords{Multicriteria optimization \and Lexicographic ordering \and Global optimization \and Dimensionality reduction \and Search information \and Computational complexity}
 \end{abstract}
@@
 \begin{equation}\label{eq:1}
-f(y) = (f_1(y), f_2(y), \dots , f_s(y)) \to min,  y \in D,
+f(y) = (f_1(y), f_2(y), \dots , f_s(y)) \to min, \; y \in D,
 \end{equation}
@@
 \begin{equation}\label{eq:2}
-D  = \{ y\in R^N: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}
+D  = \{ y\in R^N: a_i \leq y_i \leq b_i, \; 1 \leq i \leq N \}
 \end{equation}
@@
 \begin{equation}\label{eq:2l}
-|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, y',y''\in D, 1 \leq i \leq s
+|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, \; y',y''\in D,\; 1 \leq i \leq s,
 \end{equation}
@@
 This approach is also widely known as the method of successive concessions (MSC) \cite{c3,c4,c5}.
-The choice of feasible concessions $\delta_i$, $1 \leq i < s$ in the computational scheme (\ref{eq:7}) allows one to obtain any efficient decision $\widetilde{y}^* \in P_{lex}(f,D)$ and take into account the peculiarities of the MCOlex problem being solved.
+The choice of feasible concessions $\delta_i$, $1 \leq i < s$, in the computational scheme (\ref{eq:7}) allows one to obtain any efficient decision $\widetilde{y}^* \in P_{lex}(f,D)$ and take into account the peculiarities of the MCOlex problem being solved.
 At the same time, this approach makes it necessary, at each stage of scheme (\ref{eq:7}), to solve more complex global optimization problems with nonlinear constraints.
@@
 
-The extended formulation of lexicographic multicriteria optimization problems (\ref{eq:7})-(\ref{eq:8}) involves multiple solutions of multiextremal optimization problems with nonlinear constraints.
+The extended formulation of lexicographic multicriteria optimization problems (\ref{eq:7})--(\ref{eq:8}) involves multiple solutions of multiextremal optimization problems with nonlinear constraints.
 Problems of this type are computationally time-consuming and subject to the ``curse of dimensionality'' -- computational complexity increases exponentially with increasing dimensionality.
@@
 
-Adaptive selection of points $y^{k+1} \in D$ at the iterations of the global search involves the analysis of volumetric multidimensional search information $\omega_k (k \gg 1)$, which determines the high computational complexity of the decisive rules $G_k$,$k=1,2,\dots$ of global search algorithms.
-This computational complexity can be significantly simplified by reducing the optimization problems to be solved using Peano \textit{space-filling curves} or \textit{evolvents} $y(x)$, that uniquely and continuously map the interval $[0,1]$ to an $N$-dimensional hypercube D (see, e.g., \cite{c17,c18,c23}).
+Adaptive selection of points $y^{k+1} \in D$ at the iterations of the global search involves the analysis of volumetric multidimensional search information $\omega_k$ $(k \gg 1)$, which determines the high computational complexity of the decisive rules $G_k$, $k=1,2,\dots$ of global search algorithms.
+This computational complexity can be significantly simplified by reducing the optimization problems to be solved using Peano \textit{space-filling curves} or \textit{evolvents} $y(x)$, that uniquely and continuously map the interval $[0,1]$ to an $N$-dimensional hypercube $D$ (see, e.g., \cite{c17,c18,c23}).
 As a result of this reduction, the initial multidimensional multicriteria optimization problem (\ref{eq:5}) is reduced to a one-dimensional problem:
 \begin{equation}\label{eq:10}
-f(y(x)) = (f_1(y(x)), f_2(y(x)), \dots , f_s(y)(x)) \to min_{leq},  x \in [0,1].
+f(y(x)) = (f_1(y(x)), f_2(y(x)), \dots , f_s(y(x)) \to min_{leq},  x \in [0,1].
 \end{equation}
@@
 \begin{equation}\label{eq:11}
-|f_i (y(x'))-f_i (y(x';))| \leq H_i \|x'-x''\|, x',x''\in [0,1], 1 \leq i \leq s,
+|f_i (y(x'))-f_i (y(x''))| \leq H_i \|x'-x''\|, \; x',x''\in [0,1],\; 1 \leq i \leq s,
 \end{equation}
@@
 \begin{equation}\label{eq:12}
-A_k=\{(x_i, y(x_i), f_i=f(y(x_i)): 1 \leq i \leq k \},
+A_k=\{(x_i, y(x_i), f_i=f(y(x_i))): 1 \leq i \leq k \},
 \end{equation}
@@
 \begin{equation}\label{eq:13}
-\min{\{\varphi(x):g_i(x)\leq 0, 1 \leq i \leq m, x\in \Delta=[0,1]\}}
+\min{\{\varphi(x):g_i(x)\leq 0, \; 1 \leq i \leq m,\; x\in \Delta=[0,1]\}}
 \end{equation}
@@
 
-The presence of nonlinear constraints significantly complicates the complexity of solving global optimization problems; the resulting solutions must belont to a feasible domain
+The presence of nonlinear constraints significantly complicates the complexity of solving global optimization problems; the resulting solutions must belong to a feasible domain
 \begin{equation}\label{eq:14}
@@
 \begin{equation}\label{eq:18}
-A_k=\{(x_i, z_i, \nu_i, y(x_i), f_i=f( y(x_i)): 1 \leq i \leq k \},
+A_k=\{(x_i, z_i, \nu_i, y(x_i), f_i=f( y(x_i))): 1 \leq i \leq k \},
 \end{equation}
@@
 The calculation results are constantly added to the set $A_k$ from (\ref{eq:12}); in addition, into $A_k$ are also added the points $x_0=0$ and $x_{k+1}=1$ (the values $z_0$ and $z_{k+1}$ are not defined and not used) for the convenience of subsequent formulations.
-The choice of the point $x^{k+1}$, $k\geq 1$ of any next trial is determined by the following rules.
+The choice of the point $x^{k+1}$, $k\geq 1$, of any next trial is determined by the following rules.
 
@@
 
-Results of using the AGCS algorithm to solve the test problem from Fig.
-\ref{fig:1} with the parameters $r_1=r_2=r_3=2$ and $\varepsilon=10^5$are shown in Fig.
+Results of using the AGCS algorithm to solve the test problem from Fig.~\ref{fig:1} with the parameters $r_1=r_2=r_3=2$ and $\varepsilon=10^{-5}$ are shown in Fig.
 \ref{fig:2}.
@@
 	\begin{matrix}
-	\delta_j=z_j^k+\theta_j (g_j^k-z_j^k ), 0\leq \theta_j\leq 1,1\leq j < s,\\
-	g_j^k= \max{\{ g_j (x_i ) : \nu(x_i)=j,1\leq i\leq k \}}.
+	\delta_j=z_j^k+\theta_j (g_j^k-z_j^k ),\; 0\leq \theta_j\leq 1, \; 1 \leq j < s,\\
+	g_j^k= \max{\{ g_j (x_i ) : \nu(x_i)=j,\; 1\leq i\leq k \}}.
 	\end{matrix}
@@
 
-The applicability of such a one-step scheme (\ref{eq:26})-(\ref{eq:27}) instead of the scheme (\ref{eq:7}) is computationally evaluated in Section \ref{sec:4}.
+The applicability of such a one-step scheme (\ref{eq:26})--(\ref{eq:27}) instead of the scheme (\ref{eq:7}) is computationally evaluated in Section \ref{sec:4}.
 

### 2020-01-28 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
исправил форматирование таблиц

@@
 \resizebox{\columnwidth}{!}{
-\begin{tabular}{ccccccccc}
+\begin{tabular}{ccccccccccc}
 \hline
-\multicolumn{8}{c}{Search information}                                    & Reduction \\ \cline{1-8}
-\multicolumn{4}{c}{not used}       & \multicolumn{4}{c}{used}             &  in \\ \cline{1-8}
-Method     & Problem & Avg   & Avg   & Method     & Problem & Avg   & Avg   & iteration    \\
-iter.      & solved  & HV    & DU    & iter.      & solved  & HV    & DU    & number       \\ \hline
-41 710.1   & 98.0\%  & 33.43 & 0.173 & 2 407.41   & 99.3\%  & 33.34 & 0.224 & 17.3         \\ \hline
+\multicolumn{10}{c}{Search information}                                    & Reduction \\ \cline{1-9}
+\multicolumn{4}{c}{not used}         & & \multicolumn{4}{c}{used}             &  & in \\ \cline{1-4} \cline{6-9}
+Method     & Problem & Avg   & Avg   & & Method     & Problem & Avg   & Avg   &  & iteration    \\
+iter.      & solved  & HV    & DU    & & iter.      & solved  & HV    & DU    &  & number       \\ \cline{1-4} \cline{6-9} \cline{11-11}
+41 710.1   & 98.0\%  & 33.43 & 0.173 & & 2 407.41   & 99.3\%  & 33.34 & 0.224 &  & 17.3         \\ \hline
 \end{tabular}
@@
 \resizebox{\columnwidth}{!}{
-\begin{tabular}{ccccccccc}
+\begin{tabular}{ccccccccccc}
 \hline
-\multicolumn{8}{c}{Search information}                                    & Reduction \\ \cline{1-8}
-\multicolumn{4}{c}{not used}       & \multicolumn{4}{c}{used}             &  in \\ \cline{1-8}
-Method      & Problem & Avg   & Avg   & Method     & Problem & Avg   & Avg   & iteration    \\
-iter.       & solved  & HV    & DU    & iter.      & solved  & HV    & DU    & number       \\ \hline
-4 536 377.9 & 83.0\%  & 30.67 & 0.335 & 709 014.9  & 94.5\%  & 30.46 & 0.405 & 6.4         \\ \hline
+\multicolumn{10}{c}{Search information}                                        & Reduction   \\ \cline{1-9}
+\multicolumn{4}{c}{not used}       & & \multicolumn{4}{c}{used}                & &  in       \\ \cline{1-4} \cline{6-9}
+Method      & Problem & Avg   & Avg   & & Method     & Problem & Avg   & Avg   & & iteration \\
+iter.       & solved  & HV    & DU    & & iter.      & solved  & HV    & DU    & & number    \\ \cline{1-4} \cline{6-9} \cline{11-11}
+4 536 377.9 & 83.0\%  & 30.67 & 0.335 & & 709 014.9  & 94.5\%  & 30.46 & 0.405 & & 6.4       \\ \hline
 \end{tabular}

### 2020-01-28 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
Fix misprints suggested by Harley

@@
 Various methods of approximating complex constraints with the aid of simpler forms (linear, convex, etc.) are widely used.
-However, the penalty function method is most commonly used, when calculating in an unfeasble region $\Delta \setminus \Delta_g$, some ``penalty'' are added to the values of the minimized function.
+However, the penalty function method is most commonly used, when calculating in an unfeasible region $\Delta \setminus \Delta_g$, some ``penalty'' is added to the values of the minimized function.
 But the penalty function method often involves repeatedly solving the optimization problem to determine a sufficient value of the penalty coefficient.
@@
 We apply a sequential scheme to obtain the constraint values $g_i(y(x)) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$, according to which the constraint values are calculated strictly in the order of their numbers, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme).
-The number of constraints in which the values were calculated will be referred below as the index $\nu=\nu(x)$, that is
+The number of constraints in which the values were calculated will be indicated below as the index $\nu=\nu(x)$, that is
 \begin{equation}\label{eq:15}
@@
 \end{equation}
-where for the unknown a priori values $f_j^*$, $1 \leq j < s$, can be used the estimates $z_j^k$, $1 \leq j < s$, from (\ref{eq:20}) obtained from the search information $A_k$,  and for the constants $\delta_j$, $1 \leq j<s$, we can use the values normalized to the range of constraints values, that is
+where for the unknown a priori values $f_j^*$, $1 \leq j < s$, can be used the estimates $z_j^k$, $1 \leq j < s$, from (\ref{eq:20}) obtained from the search information $A_k$,  and for the constants $\delta_j$, $1 \leq j<s$, we can use the values normalized to the range of the constraints' values, that is
 \begin{equation}\label{eq:27}
@@
 
-The results from the executed experiments demonstrate that the MAGCS algorithm has a significant advantage in comparison with the considered multicriteria optimization methods of, even when solving relatively simple MCO problems.
+The results from the executed experiments demonstrate that the MAGCS algorithm has a significant advantage in comparison with the considered multicriteria optimization methods, even when solving relatively simple MCO problems.
 
@@
 As a means of control, we compared the solution points found by MAGCS and the points calculated from the Pareto boundary, taking into account the selected values of $\theta_j$, $1 \leq j < s$.
-The method parameters were set as foolows: the accuracy $\varepsilon = 0.02$ and the reliability $r_1=r_2=5.6$.
+The method parameters were set as follows: the accuracy $\varepsilon = 0.02$ and the reliability $r_1=r_2=5.6$.
 Results of the computational experiments are presented in Table 2.
@@
 The criteria of the MCO problems being solved were determined using the GKLS generator \cite{c37}.
-The method parameters were set as foolows: the accuracy $\varepsilon = 0.025$ and the reliability $r_1=r_2=5.6$.
+The method parameters were set as follows: the accuracy $\varepsilon = 0.025$ and the reliability $r_1=r_2=5.6$.
 The results of computational experiments are presented in Table 3.

### 2020-01-28 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
add two references and correct formulas

@@
 
-\title{A Computationally Efficient Approach for Solving Lexicographic Multicriteria Optimization Problems\thanks{This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''.}
+\title{A Computationally Efficient Approach for Solving Lexicographic Multicriteria Optimization Problems\thanks{This research was supported by the Russian Science Foundation, project No 16-11-10150.
+%``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''.
+}
 }
@@
 \begin{equation}\label{eq:6}
-D_1=Arg \min_{y \in D}{f_1(y)},\dots, D_i=Arg \min_{y \in D_{i-1}}{f_i(y)},1 \leq i \leq s,D_0=D,
+D_0=D,\; D_1=Arg \min_{y \in D_0}{f_1(y)},\dots, D_i=Arg \min_{y \in D_{i-1}}{f_i(y)},1 \leq i \leq s,
 \end{equation}
@@
 The extended formulation of lexicographic multicriteria optimization problems (\ref{eq:7})--(\ref{eq:8}) involves multiple solutions of multiextremal optimization problems with nonlinear constraints.
-Problems of this type are computationally time-consuming and subject to the ``curse of dimensionality'' -- computational complexity increases exponentially with increasing dimensionality.
+Problems of this type are computationally time-consuming and subject to the ``curse of dimensionality'': computational complexity increases exponentially with increasing dimensionality.
 The proposed approach for solving such problems is based on three key ideas:
 \begin{itemize}
-  \item Reduction of multidimensional problems of multiextremal optimization to problems of one-dimensional global search by using mappings based on Peano curves;
-  \item Using the index method of global optimization, which allows solving optimization problems with constraints without using penalty functions;
-  \item Repeated use of search information obtained in the process of computing, for solving multiple global optimization problems.
+  \item reduction of multidimensional problems of multiextremal optimization to problems of one-dimensional global search by using mappings based on Peano curves;
+  \item using the index method of global optimization, which allows solving optimization problems with constraints without using penalty functions;
+  \item repeated use of search information obtained in the process of computing, for solving multiple global optimization problems.
 \end{itemize}
@@
 \end{equation}
-where $k$ is the number of the global search iterations being performed, $G_k$ is the decisive rule of the applied global search algorithm, according to which the point $y^{k+1}$ of the next iteration is selected, and $\omega_k$ is the search information available at the $k$ step of the global search.
+where $k$ is the number of the global search iterations being performed, $G_k$ is the decision rule of the applied global search algorithm, according to which the point $y^{k+1}$ of the next iteration is selected, and $\omega_k$ is the search information available at the $k$ step of the global search.
 
-Adaptive selection of points $y^{k+1} \in D$ at the iterations of the global search involves the analysis of volumetric multidimensional search information $\omega_k$ $(k \gg 1)$, which determines the high computational complexity of the decisive rules $G_k$, $k=1,2,\dots$ of global search algorithms.
+Adaptive selection of points $y^{k+1} \in D$ at the iterations of the global search involves the analysis of volumetric multidimensional search information $\omega_k$, $k \gg 1$, which determines the high computational complexity of the decision rules $G_k$, $k=1,2,\dots$, of global search algorithms.
 This computational complexity can be significantly simplified by reducing the optimization problems to be solved using Peano \textit{space-filling curves} or \textit{evolvents} $y(x)$, that uniquely and continuously map the interval $[0,1]$ to an $N$-dimensional hypercube $D$ (see, e.g., \cite{c17,c18,c23}).
@@
 \subsection{An efficient method for solving global optimization problems with 
-nonlinear constraints}
+non-convex constraints}
 
-The problems solved in the framework of the computational scheme (\ref{eq:7}) are one-dimensional global optimization problems with nonlinear constraints
+The problems solved in the framework of the computational scheme (\ref{eq:7}) are one-dimensional global optimization problems with non-convex constraints
 \begin{equation}\label{eq:13}
@@
 
-The presence of nonlinear constraints significantly complicates the complexity of solving global optimization problems; the resulting solutions must belong to a feasible domain
+The presence of non-convex constraints significantly complicates the complexity of solving global optimization problems; the resulting solutions must belong to a feasible domain
 \begin{equation}\label{eq:14}
@@
 But the penalty function method often involves repeatedly solving the optimization problem to determine a sufficient value of the penalty coefficient.
-More complete information on methods for solving optimization problems with constraints can be obtained, for example, in monographs \cite{c18,c26,c27}.
+More complete information on methods for solving optimization problems with constraints can be obtained, for example, in monographs \cite{c18,c26,c27,c28,c29}.
 
@@
 
-We apply a sequential scheme to obtain the constraint values $g_i(y(x)) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$, according to which the constraint values are calculated strictly in the order of their numbers, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme).
+We apply a sequential scheme to obtain the constraint values $g_i(y(x))$, $1 \leq i \leq m$, and the objective function $\varphi(x)$, according to which the constraint values are calculated strictly in the order of their numbers, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme).
 The number of constraints in which the values were calculated will be indicated below as the index $\nu=\nu(x)$, that is
 \begin{equation}\label{eq:15}
-\nu(x) : g_\nu(x)>0, g_j(x)\leq 0, 1 \leq j\leq \nu - 1, 1\leq \nu = \nu(x) \leq m
+\nu(x) : g_\nu(x)>0,\; g_j(x)\leq 0,\; 1 \leq j\leq \nu - 1,\; 1\leq \nu = \nu(x) \leq m
 \end{equation}
@@
 
-The function $F(x)$ consists of separate fragments of the constraints $g_i(x) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$.
+The function $F(x)$ consists of separate fragments of the constraints $g_i(x)$, $1 \leq i \leq m$, and the objective function $\varphi(x)$.
 For the homogeneity of the components, the function $F(x)$ is then converted to the form
@@
 \begin{equation}\label{eq:21}
-h^k_l=\max{\{|z_i-z_j|/\sqrt[N]{x_i-x_j},\nu_i=\nu_j=l,i>j,1\leq l\leq M^k\}}
+h^k_l=\max{\{|z_i-z_j|/\sqrt[N]{x_i-x_j},\; \nu_i=\nu_j=l,\; i>j, \;1\leq l\leq M^k\}}.
 \end{equation}
@@
 If $h_l^k$ in (\ref{eq:21}) turns out to be indefinite or equal to zero, then $h_l^k=1$ is taken.
-For the initial values of these quantities, $M^0 = H_l^0=1$, $z_l^0=0$, $1 \leq i \leq m+1$ can be taken.
+For the initial values of these quantities, $M^0 = H_l^0=1$, $z_l^0=0$, $1 \leq i \leq m+1$, can be taken.
 
-The general computational scheme of the global search algorithm for multiextremal optimization problems with nonlinear constraints \footnote{This method is also known as the index method -- see \cite{c18}.} (AGCS) is as follows \cite{c18}.
+The general computational scheme of the global search algorithm for multiextremal optimization problems with non-convex constraints \footnote{This method is also known as the index method -- see \cite{c18}.} (AGCS) is as follows \cite{c18}.
 
@@
 
-As noted earlier, the MCOlex problem at each stage of the successive concessions method (\ref{eq:7}) is a global optimization problem with nonlinear constraints (\ref{eq:13}).
+As noted earlier, the MCOlex problem at each stage of the successive concessions method (\ref{eq:7}) is a global optimization problem with non-convex constraints (\ref{eq:13}).
 Generally, the solution to such problems must start at the beginning, and multiple solutions of such global optimization problems may require extensive calculations.
@@
 
-The AGCS algorithm, supplemented by the ability to use search information to solve multiple MCOlex  problems, will be referred to below as the global search algorithm for multistage solutions of a set of multiextremal optimization problems with nonlinear constraints (MAGCS).
+The AGCS algorithm, supplemented by the ability to use search information to solve multiple MCOlex  problems, will be referred to below as the global search algorithm for multistage solutions of a set of multiextremal optimization problems with non-convex constraints (MAGCS).
 
@@
 
-The global optimization algorithms used within the framework of the proposed approach have proven to be efficient in the process of conducting test computational experiments and have been widely used in solving practical optimization problems (see, e.
-g., \cite{c33,c34}).
+The global optimization algorithms used within the framework of the proposed approach have proven to be efficient in the process of conducting test computational experiments and have been widely used in solving practical optimization problems (see, e.g., \cite{c33,c34}).
 The results of computational experiments in solving multicriteria optimization problems are given below.
@@
 
-Overcoming the enormous computational complexity of addressing the formulated new class of MCOlex problems is ensured by solving a sequence of global optimization problems with nonlinear constraints using efficient information-statistical methods of global optimization by using an original index constraint accounting scheme instead of the commonly used penalty functions.
+Overcoming the enormous computational complexity of addressing the formulated new class of MCOlex problems is ensured by solving a sequence of global optimization problems with non-convex constraints using efficient information-statistical methods of global optimization by using an original index constraint accounting scheme instead of the commonly used penalty functions.
 A core element in this approach is the ability to use all the search information obtained in the computation process, utilizing the multistage solution of MCOlex problems.

### 2020-01-28 2019/NUMTA_Barkalov/Special Issure/article.tex
New results for Simple_4 class
Last corrections before submitting

@@
 \begin{equation}\label{lipschitz_condition}
-	\left|g_j(y')-g_j (y'')\right| \leq L_j \|y'-y''\|, \; y', y''\in D, \; 1\leq j \leq m+1.
+	\left|g_j(y')-g_j (y'')\right| \leq L_j \|y'-y''\|, \; y', y''\in D, \; 1\leq j \leq m+1,
 \end{equation}
@@
 
-	The existence of the $\varepsilon$-reserved solutions of problem (\ref{problem}) can be interpreted as some analog of the regularity condition in classical nonlinear programming problems.
+	The existence of the $\varepsilon$-reserved solutions of problem (\ref{problem}) can be interpreted as some analog of the regularity conditions in classical nonlinear programming problems.
 The applied role of this condition should be also noted.
@@
 	Thus, a search trial at some point $x^k \in [0,1]$ performed at the $k^{th}$ iteration of the algorithm will consist of the following sequence of operations:
-\begin{enumerate} 
-  \item Determine the image $y^k=y(x^k )$ with the use of the evolvent $y(x)$. 
-  \item Compute the values $g_1 (y^k ), \ldots ,g_{\nu} (y^k )$ where $\nu = \nu(x)$ is the index of the point.
-	\item Fix the result of the trial as three values 
-\end{enumerate}
-
+\begin{itemize} 
+  \item determine the image $y^k=y(x^k )$ with the use of the evolvent $y(x)$. 
+  \item compute the values $g_1 (y^k ), \ldots ,g_{\nu} (y^k )$, where $\nu = \nu(x^k)$ is the index of the point $x^k$.
+	\item fix the result of the trial as three values 
 \begin{equation}\label{trial_result}
-	x^k, \; z^k = g_{\nu}\left( y(x^k) \right), \; \nu = \nu(x^k),
+	x^k, \; z^k = g_{\nu}\left( y(x^k) \right), \; \nu = \nu(x^k).
 \end{equation}
+\end{itemize}
 
@@
 when $x^k \in Q_{m+1}$, the trial involves checking all constraints and computing the objective function.
-Here, the value of the index is accepted to be equal to the quantity $\nu = m+1$.
+Here, the value of the index $\nu$ is accepted to be equal to $m+1$.
 
@@
      0, & \nu \leq m,\\
-     \frac{\varphi^\ast}{K_{m+1}}, & \nu = m + 1.
+     \frac{\varphi^\ast}{K_{m+1}}, & \nu = m + 1,
    \end{array}
@@
 	The index-based global search algorithm under consideration for solving the problem (\ref{problem}) (here it is given according to \cite{Strongin2000}) involves the construction of a sequence of points $x^k \in (0,1)$, at which the search trials are conducted with the results of the type (\ref{trial_result}).
-The set of triplets $\{x^k, z^k=g_{\nu}\big(y(x^k)\big), \nu=\nu(x^k)\}, 1 \leq k \leq n,$  makes up the search information accumulated by the method after performing $n$ steps.
-The operation of the index-based global search algorithm is defined by the following rules.
+The set of triplets
+$\{x^k,\; z^k=g_{\nu}\big(y(x^k)\big),\; \nu=\nu(x^k)\},\; 1 \leq k \leq n,$
+makes up the search information accumulated by the method after performing $n$ steps. The operation of the index-based global search algorithm is defined by the following rules.
 
@@
 \\
-\textbf{Theorem}. Assume that the following conditions are satisfied:
+\textbf{Theorem 1}. Assume that the following conditions are satisfied:
 \begin{enumerate} 
@@
 
-	At the same time, the index scheme of accounting for the constraints also demonstrated its advantages (in the number of computed values of the problem functions) over the traditional penalty functions method in solving a wide range of model problems (see the results of the experimental comparison in \cite{Barkalov2017_1, Barkalov2017_2}.)
+	At the same time, the index scheme of accounting for the constraints also demonstrated its advantages (in the number of computed values of the problem functions) over the traditional penalty functions method in solving a wide range of test problems (see the results of the experimental comparison in \cite{Barkalov2017_1, Barkalov2017_2}.)
 
@@
 
-	According to Theorem 1 from Section 3, the sequence of trial points $\{y^k\}$ will converge to a global minimizer $y^\ast$ if the condition (\ref{theorem_inequalities}) is satisfied.
+	According to Theorem 1, the sequence of trial points $\{y^k\}$ will converge to a point $\bar y \in Y_\epsilon$ if the condition (\ref{theorem_inequalities}) is satisfied.
 So far, an appropriate choice of the parameters $r_{\nu}$ allows using the values
@@
 
-	Compliance with condition (\ref{theorem_inequalities}) will be achieved when choosing a high value of the parameters $r_{\nu}$.
+	Compliance with condition (\ref{theorem_inequalities}) will be achieved when choosing a high values of the parameters $r_{\nu}$.
 However, in this case the constant will be overestimated, and the method will perform a large number of trials until the stopping condition is fulfilled.
@@
 
-	Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1,$ 
+	Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$: 
 \begin{flushleft}
 	\begin{itemize}[\labelitemii]
-		\item Compute the global characteristics $R_{glob}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{glob}$.
-		\item Compute the local characteristics $R_{loc}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{loc}$.
-		\item Determine the characteristic $R(i)$ as
+		\item compute the global characteristics $R_{glob}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{glob}$;
+		\item compute the local characteristics $R_{loc}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{loc}$;
+		\item determine the characteristic $R(i)$ as
 \begin{equation}\label{R_max}
@@
 \end{equation}
-where $\rho=\Big(\frac{1-1/r_{M}^{glob}}{1-1/r_{M}^{loc}}\Big)^2$, if $\nu(x_{i-1})=\nu(x_i)=\nu$, and $\rho=1$, if $\nu(x_{i-1}) \neq \nu(x_i)$.
+where $\rho=\Big(\frac{1-1/r_{\nu}^{glob}}{1-1/r_{\nu}^{loc}}\Big)^2$ if $\nu(x_{i-1})=\nu(x_i)=\nu$, and $\rho=1$ if $\nu(x_{i-1}) \neq \nu(x_i)$.
 	\end{itemize}
@@
 \end{eqnarray}
-The feasible domain of this problem is not connected and consists of three subdomains.
+The feasible domain of this problem is non-connected and consists of three subdomains.
 The global minimum $\varphi(y^\ast)=-1.489$ is achieved at the point $y^\ast =(0.942, 0.944)$ (the values are given with the accuracy of $10^{-3}$).
@@
 For both methods, the evolvent with the parameter $m=10$ was used, the accuracy of the search for the solution $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ from (\ref{epsilon_nu}).
-For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$ were used.
+For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$, were used.
 The same values were used for the IA-DL method as the parameters $r_{\nu}^{glob}, 1 \leq \nu \leq 4$.
@@
 Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem 1 are fulfilled.
-Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1, 1\leq \nu \leq m+1,$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e.
+Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1,\; 1\leq \nu \leq m+1,$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e.
 satisfies the condition
@@
 
-	Since the limit point $\bar y = y(\bar x)$ belongs to the set $Y_{\epsilon}$, then
+	Since the point $\bar y = y(\bar x)$ belongs to the set $Y_{\epsilon}$, then
 $$
@@
 	&\geq& \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	&=& \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \geq 0.
+	&=& \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
 	\nonumber
@@
 			\hline
-			IA-DL, $r_{\nu}^{loc}=1.4$ & 316 & 7403 & 44304 & 296357 \\
+			IA-DL, $r_{\nu}^{loc}=1.4$ & 316 & 7403 & 40419 & 296357 \\
 			\hline
-			IA-DL, $r_{\nu}^{loc}=2.0$ & 350 & 8990(1) & 47133 & 360497 \\
+			IA-DL, $r_{\nu}^{loc}=2.0$ & 350 & 8990(1) & 41230 & 360497 \\
 			\hline
-			IA & 439 & 11804(1) & 42837 & 404660(2) \\
+			IA & 439 & 11804(1) & 41973(3) & 404660(2) \\
 			\hline

### 2020-01-28 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
мелкие правки: nonliner -> non-convex часть текста сделан курсивом

@@
 
-Among the approaches which have been developed for solving MCO problems, lexicographic optimization methods can be singled out, when some ordering of criteria by importance is implemented and optimization is carried out based on their ordering \cite{c3}.
-Iterative methods represent another approach \cite{c6,c10}, when the decision maker is actively involved in the process of selecting decisions.
+Among the approaches which have been developed for solving MCO problems, \textit{lexicographic optimization} methods can be singled out, when some ordering of criteria by importance is implemented and optimization is carried out based on their ordering \cite{c3}.
+\textit{Iterative methods} represent another approach \cite{c6,c10}, when the decision maker is actively involved in the process of selecting decisions.
 An actively appropriated approach consists in developing and applying evolutionary algorithms based on the simulation of certain natural phenomena to solve MCO problems \cite{c10,c11,c12,c13}.
-Among the widely used endeavors in solving MCO problems is the scalarization approach, which applies methods involving the convolution of criteria to a single scalar criterion (see, e.g.
+Among the widely used endeavors in solving MCO problems is the \textit{scalarization} approach, which applies methods involving the convolution of criteria to a single scalar criterion (see, e.g.
 \cite{c2,c14}).
@@
 This work is devoted to solving lexicographic multicriteria optimization (MCOlex) problems that arise when designing complex technical objects and systems.
-In such applications, efficiency criteria may have a complex multiextremal form, and calculating criteria values may require large amounts of computations.
-In such conditions, finding even one efficient decision requires significant calculations, while determining several (or a whole set) of efficient decisions becomes an issue of great computational complexity.
+In such applications, efficiency criteria may have a complex \textit{multiextremal} form, and calculating criteria values may require \textit{large amounts of computations}.
+In such conditions, finding even one efficient decision requires significant calculations, while determining several (or a whole set) of efficient decisions becomes an issue of \textit{great computational complexity}.
 
@@
 This paper proposes a solution to the time-consuming class of MCOlex problems by using an approach based on the following main points.
-First of all, the problem is reduced to solving a sequence of global optimization problems with nonlinear constraints (GCO) \cite{c2,c14}.
+First of all, the problem is reduced to solving a sequence of global optimization problems with non-convex constraints (GCO) \cite{c2,c14}.
 Then, efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization are applied for solving the GCO problems \cite{c17,c18}.
@@
 The further structure of the paper is as follows.
-Section \ref{sec:2} presents a new class of optimization problems -- multistage lexicographic multicriteria optimization (MMCOlex) problems -- whose solution is reduced to solving a sequence of global optimization problems with nonlinear constraints.
+Section \ref{sec:2} presents a new class of optimization problems -- multistage lexicographic multicriteria optimization (MMCOlex) problems -- whose solution is reduced to solving a sequence of global optimization problems with non-convex constraints.
 Section \ref{sec:3} presents the basics of the approach that has been developed: reducing multidimensional MCOlex problems to one-dimensional global optimization problems, applying efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization, and reusing the search information obtained in the process of calculations.
@@
 The choice of feasible concessions $\delta_i$, $1 \leq i < s$, in the computational scheme (\ref{eq:7}) allows one to obtain any efficient decision $\widetilde{y}^* \in P_{lex}(f,D)$ and take into account the peculiarities of the MCOlex problem being solved.
-At the same time, this approach makes it necessary, at each stage of scheme (\ref{eq:7}), to solve more complex global optimization problems with nonlinear constraints.
+At the same time, this approach makes it necessary, at each stage of scheme (\ref{eq:7}), to solve more complex global optimization problems with non-convex constraints.
 
@@
 
-The extended formulation of lexicographic multicriteria optimization problems (\ref{eq:7})--(\ref{eq:8}) involves multiple solutions of multiextremal optimization problems with nonlinear constraints.
+The extended formulation of lexicographic multicriteria optimization problems (\ref{eq:7})--(\ref{eq:8}) involves multiple solutions of multiextremal optimization problems with non-convex constraints.
 Problems of this type are computationally time-consuming and subject to the ``curse of dimensionality'': computational complexity increases exponentially with increasing dimensionality.
@@
 
-The general computational scheme of the global search algorithm for multiextremal optimization problems with non-convex constraints \footnote{This method is also known as the index method -- see \cite{c18}.} (AGCS) is as follows \cite{c18}.
+The general computational scheme of the \textit{global search algorithm} for multiextremal optimization problems with non-convex constraints \footnote{This method is also known as the index method -- see \cite{c18}.} (AGCS) is as follows \cite{c18}.
 

### 2020-01-28 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
delete reference [18]

@@
 
-The general computational scheme of the \textit{global search algorithm} for multiextremal optimization problems with non-convex constraints \footnote{This method is also known as the index method -- see \cite{c18}.} (AGCS) is as follows \cite{c18}.
+The general computational scheme of the \textit{global search algorithm} for multiextremal optimization problems with non-convex constraints \footnote{This method is also known as the index method.} (AGCS) is as follows \cite{c18}.
 

### 2020-01-31 2019/NUMTA_Barkalov/SoftComputing/article.tex
Add SoftComputing

--- 2020-01-28 2019/NUMTA_Barkalov/Special Issure/article.tex
@@
-\documentclass[preprint]{elsarticle}
-
+% First comes an example EPS file -- just ignore it and
+% proceed on the \documentclass line
+% your LaTeX will extract the file if required
+\begin{filecontents*}{example.eps}
+gsave
+newpath
+  20 20 moveto
+  20 220 lineto
+  220 220 lineto
+  220 20 lineto
+closepath
+2 setlinewidth
+gsave
+  .4 setgray fill
+grestore
+stroke
+grestore
+\end{filecontents*}
+%
+\RequirePackage{fix-cm}
+%
+%\documentclass{svjour3}                     % onecolumn (standard format)
+%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
+%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
+\documentclass[twocolumn]{svjour3}          % twocolumn
+%
+\smartqed  % flush right qed marks, e.g. at end of proof
+%
+\usepackage{graphicx}
 \usepackage{lineno,hyperref}
-%\usepackage[T1]{fontenc}
-\modulolinenumbers[5]
-
-%\journal{Journal of \LaTeX\ Templates}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%% Elsevier bibliography styles
-%%%%%%%%%%%%%%%%%%%%%%%
-%% To change the style, put a % in front of the second line of the current style and
-%% remove the % from the second line of the style you would like to use.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Numbered
-%\bibliographystyle{model1-num-names}
-
-%% Numbered without titles
-%\bibliographystyle{model1a-num-names}
-
-%% Harvard
-%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
-
-%% Vancouver numbered
-%\usepackage{numcompress}\bibliographystyle{model3-num-names}
-
-%% Vancouver name/year
-%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
-
-%% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
-
-%% AMA style
-%\usepackage{numcompress}\bibliographystyle{model6-num-names}
-
-%% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
-%%%%%%%%%%%%%%%%%%%%%%%
-
+%
+% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
+%
+% insert here the call for the packages your document requires
+%\usepackage{latexsym}
+% etc.
+%
+% please place your own definitions here and don't use \def but
+% \newcommand{}{}
+%
+% Insert the name of "your journal" with
+% \journalname{myjournal}
+%
 \begin{document}
 
-\begin{frontmatter}
+\title{Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints%\thanks{Grants or other notes
+%about the article that should go on the front page should be
+%placed here. General acknowledgments should be placed at the end of the article.}
+}
+%\subtitle{Do you have a subtitle?\\ If so, write it here}
 
-\title{Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints}
+%\titlerunning{Short form of title}        % if too long for running head
 
-%% Group authors per affiliation:
-\author{Roman Strongin}
-\author{Konstantin Barkalov}
-\author{Semen Bevzuk}
-%\address[mymainaddress]{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia}
-%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
-%\cortext[mycorrespondingauthor]{Corresponding author}
-%\ead{support@elsevier.com}
-\address{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia}
+\author{Roman Strongin        \and
+        Konstantin Barkalov   \and
+				Semen Bevzuk
+}
+
+%\authorrunning{Short form of author list} % if too long for running head
+
+\institute{F. Author \at
+              first address \\
+              Tel.: +123-45-678910\\
+              Fax: +123-45-678910\\
+              \email{fauthor@example.com}           %  \\
+%             \emph{Present address:} of F. Author  %  if needed
+           \and
+           S. Author \at
+              second address
+}
+
+\date{Received: date / Accepted: date}
+
+\maketitle
 
 \begin{abstract}
+%Insert your abstract here. Include keywords, PACS and mathematical
+%subject classification numbers as needed.
 This paper considers the constrained global optimization problems, in which the functions are of the “black box” type and satisfy the Lipschitz condition.
@@
 The results of the numerical experiments on a series of multiextremal problems with non-convex constraints demonstrating the efficiency of the proposed scheme of dual Lipschitz constant estimates are presented.
+\keywords{Global optimization \and Multiextremal problems \and Non-convex constraints \and Lipschitz constant estimates \and Dimensionality reduction \and Numerical methods}
+% \PACS{PACS code1 \and PACS code2 \and more}
+% \subclass{MSC code1 \and MSC code2 \and more}
 \end{abstract}
 
-\begin{keyword}
-Global optimization, Multiextremal problems, Non-convex constraints, Lipschitz constant estimates, Dimensionality reduction, Numerical methods
-\end{keyword}
-
-\end{frontmatter}
-
-\linenumbers
-
 \section{Introduction}
+\label{intro}
 In this paper, we consider constrained global optimization problems and the numerical methods for solving problems of this type.
@@
 	The methods for solving the Lipschitzian optimization problems constitute a very important area in the development of the global optimization methods and are the objects of investigations of many researchers.
-The first algorithms of this class were proposed as early as in the 1970s \cite{Evtushenko1971, Piyavskii1972, Shubert1972, Strongin1970}; since that time, this area has been continuing to develop extensively (see, for example \cite{Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Jones2009}).
+The first algorithms of this class were proposed as early as in the 1970s \cite{Evtushenko1971,Piyavskii1972,Shubert1972,Strongin1970}; since that time, this area has been continuing to develop extensively (see, for example \cite{Evtushenko2009,Evtushenko2013,Strongin2000,Sergeyev2013,Jones2009}).
 Note that these global optimization algorithms are, as a rule, oriented towards solving either  unconstrained optimization problems
-	\cite{Paulavicius2014, Sergeyev2015, Pinter1996, Jones1993, Gablonsky2001}
-or problems with constraints of a simple structure \cite{Vaz2009, Paulavicius2016}.
-The solving of the problems with complex non-convex constraints is usually performed with the use of the penalty functions method \cite{Stripinis2019, Pillo2012, Pillo2016}, which has a number of drawbacks.
+	\cite{Paulavicius2014,Sergeyev2015,Pinter1996,Jones1993,Gablonsky2001}
+or problems with constraints of a simple structure \cite{Vaz2009,Paulavicius2016}.
+The solving of the problems with complex non-convex constraints is usually performed with the use of the penalty functions method \cite{Stripinis2019,Pillo2012,Pillo2016}, which has a number of drawbacks.
 In particular, this method cannot be applied to the problems with the partially defined functions (i.e.
@@
 This allows solving the problems with partially defined functions.
-The experimental comparison of the index method of accounting for the constraints and the penalty function method demonstrated the advantages of the index method in terms of the number of iterations and the number of computations of the problem functions \cite{Barkalov2017_1, Barkalov2017_2}.
+The experimental comparison of the index method of accounting for the constraints and the penalty function method demonstrated the advantages of the index method in terms of the number of iterations and the number of computations of the problem functions \cite{Barkalov2017_1,Barkalov2017_2}.
 
 	Note that the values of the Lipschitz constant for the problem functions are usually not known a priori, which  makes their estimation one of the key problems in the construction of the Lipschitzian optimization methods.
-The methods utilizing the values of the Lipschitz constant predefined a priori (see, for example, \cite{Piyavskii1972, Shubert1972, Wood1991, Meewella1988, Mladineo1986}) are important in the theoretical aspect, but it is difficult to use such methods in applied problems (in which the information of the Lipschitz constant values is absent).
+The methods utilizing the values of the Lipschitz constant predefined a priori (see, for example, \cite{Piyavskii1972,Shubert1972,Wood1991,Meewella1988,Mladineo1986}) are important in the theoretical aspect, but it is difficult to use such methods in applied problems (in which the information of the Lipschitz constant values is absent).
 
@@
 		Several typical methods for adaptive estimation of the Lipschitz constant according to the results of the performed search trials are known:
-\begin{itemize}[\labelitemii]
+\begin{itemize}
 \item global estimate of the constant $L$ in the whole search domain $D$ \cite{Strongin2000,Pinter1996,Horst1996}.
-  \item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ \cite{Kvasov2003, Sergeyev2010, Sergeyev2016}.
-	\item choice of the estimates of the constant $L$ from some set of possible values \cite{Jones2009,Jones1993, Gablonsky2001,Sergeyev2006}.
+\item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ \cite{Kvasov2003,Sergeyev2010,Sergeyev2016}.
+\item choice of the estimates of the constant $L$ from some set of possible values \cite{Jones2009,Jones1993,Gablonsky2001,Sergeyev2006}.
 \end{itemize}
@@
 \section{Index Method of Accounting for the Constraints and Dimensionality Reduction}
+\label{sec:1}
 	We consider the global optimization problems of the form
@@
 \end{equation}
-The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000, Sergeyev2013}.
+The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000,Sergeyev2013}.
 Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$, where the number $m>1$ is a parameter (the \textit{density} of the evolvent).
@@
 For  illustration, Fig.
-\ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
-\begin{figure}%[h]
+\ref{fig:evolvent} (a), (b) shows the projections of the interval $[0,1]$ for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
+\begin{figure}
 	\begin{minipage}[h]{0.45\linewidth}
@@
 
-	The dimensionality reduction scheme considered here matches a multidimensional problem with Lipschitzian objective function and Lipschitzian constraints to a one-dimensional problem, in which the corresponding functions satisfy the uniform H\"{o}lder condition (see \cite{Strongin2000, Sergeyev2013})
+	The dimensionality reduction scheme considered here matches a multidimensional problem with Lipschitzian objective function and Lipschitzian constraints to a one-dimensional problem, in which the corresponding functions satisfy the uniform H\"{o}lder condition (see \cite{Strongin2000,Sergeyev2013})
 $$
@@
 \section{Index-Based Global Search Algorithm}
+\label{sec:2}
 	The index-based global search algorithm under consideration for solving the problem (\ref{problem}) (here it is given according to \cite{Strongin2000}) involves the construction of a sequence of points $x^k \in (0,1)$, at which the search trials are conducted with the results of the type (\ref{trial_result}).
@@
 	R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
-\end{equation} \\
+\end{equation} 
 Step 7. Execute the next trial at the middle point of the interval $(x_{t-1}, x_t)$ if the indices of the points $x_{t-1}, x_t$ are not the same, i.e.,
@@
 
-	At the same time, the index scheme of accounting for the constraints also demonstrated its advantages (in the number of computed values of the problem functions) over the traditional penalty functions method in solving a wide range of test problems (see the results of the experimental comparison in \cite{Barkalov2017_1, Barkalov2017_2}.)
+	At the same time, the index scheme of accounting for the constraints also demonstrated its advantages (in the number of computed values of the problem functions) over the traditional penalty functions method in solving a wide range of test problems (see the results of the experimental comparison in \cite{Barkalov2017_1,Barkalov2017_2}.)
 
 \section{Index Algorithm with Dual Lipschitz Constant Estimates}
+\label{sec:3}
 	The index global search algorithm described in the previous section (hereafter referred to as IA) is intended for solving  the multiextremal problems, in which the objective function and the constraints satisfy the Lipschitz condition.
@@
 \begin{flushleft}
-	\begin{itemize}[\labelitemii]
+	\begin{itemize}
 		\item compute the global characteristics $R_{glob}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{glob}$;
@@
 	&=& \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-%	&=& \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
 	&\geq& \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
@@
 \section{Numerical experiments}
+\label{sec:4}
 To demonstrate the efficiency of the proposed algorithm, let us consider the results of solving several series of multidimensional constrained global optimization problems.
 In order to construct the problems, the functions generated by the GKLS generator \cite{Gaviano2003} were used.
-This generator of multiextremal functions is often used for investigating global optimization algorithms \cite{Paulavicius2014, Sergeyev2015,Barkalov2018}.
+This generator of multiextremal functions is often used for investigating global optimization algorithms \cite{Paulavicius2014,Sergeyev2015,Barkalov2018}.
 According to \cite{Gaviano2003}, we used simpler and more complex functions, which are  further denoted as ``Simple'' and ``Hard'', respectively.
@@
 
-	In \cite{Gergel2017_2, Gergel2019}, an approach was proposed that allows constructing the constrained global optimization problems using certain classes of multiextremal functions (in particular, the GKLS functions), in which:
+	In \cite{Gergel2017_2,Gergel2019}, an approach was proposed that allows constructing the constrained global optimization problems using certain classes of multiextremal functions (in particular, the GKLS functions), in which:
 \begin{itemize}
@@
 	\centering
+	\caption{Average number of iterations performed by the methods on “Simple” class}
 		\begin{tabular}{ccccc}
@@
 		\end{tabular}
-	\caption{Average number of iterations performed by the methods on “Simple” class}
 	\label{tab:average_number_of_iterations_simple}
@@
 	\centering
+	\caption{Average number of iterations performed by the methods on “Hard” class}
 		\begin{tabular}{ccccc}
@@
 		\end{tabular}
-	\caption{Average number of iterations performed by the methods on “Hard” class}
 	\label{tab:average_number_of_iterations_hard}
@@
 \section{Conclusion}
+\label{sec:5}
 	In this paper, a method for solving the Lipschitz optimization problems, which constitute an important class of the global optimization problems, was considered.
@@
 
-\section*{Acknowledgments}
+\section{Section title}
+\label{sec:10}
+Text with citations \cite{RefB} and \cite{RefJ}.
+\subsection{Subsection title}
+\label{sec:20}
+as required. Don't forget to give each section
+and subsection a unique label (see Sect.~\ref{sec:1}).
+\paragraph{Paragraph headings} Use paragraph headings as needed.
+\begin{equation}
+a^2+b^2=c^2
+\end{equation}
+
+% For one-column wide figures use
+\begin{figure}
+% Use the relevant command to insert your figure file.
+% For example, with the graphicx package use
+  \includegraphics{example.eps}
+% figure caption is below the figure
+\caption{Please write your figure caption here}
+\label{fig:1}       % Give a unique label
+\end{figure}
+%
+% For two-column wide figures use
+\begin{figure*}
+% Use the relevant command to insert your figure file.
+% For example, with the graphicx package use
+  \includegraphics[width=0.75\textwidth]{example.eps}
+% figure caption is below the figure
+\caption{Please write your figure caption here}
+\label{fig:2}       % Give a unique label
+\end{figure*}
+%
+% For tables use
+\begin{table}
+% table caption is above the table
+\caption{Please write your table caption here}
+\label{tab:1}       % Give a unique label
+% For LaTeX tables use
+\begin{tabular}{lll}
+\hline\noalign{\smallskip}
+first & second & third  \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+number & number & number \\
+number & number & number \\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
+\begin{acknowledgements}
 This research was supported by the Russian Science Foundation, project No. 16-11-10150.
+\end{acknowledgements}
 
+% Authors must disclose all relationships or interests that 
+% could have direct or potential influence or impart bias on 
+% the work: 
+%
+% \section*{Conflict of interest}
+%
+% The authors declare that they have no conflict of interest.
+
+% BibTeX users please use one of
+%\bibliographystyle{spbasic}      % basic style, author-year citations
+%\bibliographystyle{spmpsci}      % mathematics and physical sciences
+%\bibliographystyle{spphys}       % APS-like style for physics
+%\bibliography{}   % name your BibTeX data base
+\bibliographystyle{spmpsci}
 \bibliography{bibliography}

### 2020-01-31 2019/NUMTA_Barkalov/SoftComputing/article.tex
Fixed several formulas

@@
-% First comes an example EPS file -- just ignore it and
-% proceed on the \documentclass line
-% your LaTeX will extract the file if required
-\begin{filecontents*}{example.eps}
-gsave
-newpath
-  20 20 moveto
-  20 220 lineto
-  220 220 lineto
-  220 20 lineto
-closepath
-2 setlinewidth
-gsave
-  .4 setgray fill
-grestore
-stroke
-grestore
-\end{filecontents*}
-%
 \RequirePackage{fix-cm}
@@
 \usepackage{graphicx}
-\usepackage{lineno,hyperref}
+\usepackage{lineno,hyperref,amsmath}
 %
@@
 	It is supposed that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_{j}(y), 1 \leq j \leq m$, of the constraints satisfy the Lipschitz condition
-\begin{equation}\label{lipschitz_condition}
-	\left|g_j(y')-g_j (y'')\right| \leq L_j \|y'-y''\|, \; y', y''\in D, \; 1\leq j \leq m+1,
-\end{equation}
+\begin{align}\label{lipschitz_condition}
+	\left|g_j(y')-g_j (y'')\right| \leq L_j \|y'-y''\|, \nonumber \\ 
+	y', y''\in D, \; 1\leq j \leq m+1,
+\end{align}
 with a priori unknown constants $L_j, 1\leq j \leq m+1$, and may be multiextremal.
@@
 	Along with the exact solution $y^\ast$ of the problem (\ref{problem}), we will consider also the $\varepsilon$-reserved solution $y_{\epsilon}$, for which 
-\begin{equation}\label{epsilon_reserved_solution}
-	\varphi(y_{\epsilon})=\min{\left\{\varphi(y):y \in D, \; g_j(y)\leq -{\epsilon}_j, \; 1 \leq j \leq m\right\}},
-\end{equation}
+\begin{align}\label{epsilon_reserved_solution}
+	\varphi(y_{\epsilon}) &= \nonumber \\
+	&\min{\left\{\varphi(y):y \in D, g_j(y)\leq -{\epsilon}_j, 1 \leq j \leq m\right\}},
+\end{align}
 and the set of solutions $Y_{\epsilon}$, which are not worse than the $\varepsilon$-reserved solution in the objective function value, i.e.
-\begin{equation}\label{Y_epsilon}
-	Y_{\epsilon}=\left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
-\end{equation}
+\begin{align}\label{Y_epsilon}
+	Y_{\epsilon} &= \nonumber \\
+	&\left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
+\end{align}
 Fig. \ref{fig:eps_reserved_solution} illustrates the concepts introduced.
@@
 It means that they are defined and computable only in the sets
-\begin{equation}\label{D_sets}
-	D_1 = D, \; D_{j+1} = \left\{ y \in D_j: \; g_j(y) \leq 0 \right\}, \; 1 \leq j \leq m.
-\end{equation}
+\begin{align}\label{D_sets}
+	D_1 = D, \; D_{j+1} = \left\{ y \in D_j: \; g_j(y) \leq 0 \right\}, \nonumber \\
+	1 \leq j \leq m.
+\end{align}
 This situation is typical for the optimal design problems when some conditions for the functioning of the modeled object are not met and its other characteristics appear to be undefined.
@@
 Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
-\begin{equation}\label{one_dimensional_problem}
-	\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
-\end{equation}
+\begin{align}\label{one_dimensional_problem}
+	\varphi(y(x^\ast)) =  
+	\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
+\end{align}
 The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000,Sergeyev2013}.
@@
 	The dimensionality reduction scheme considered here matches a multidimensional problem with Lipschitzian objective function and Lipschitzian constraints to a one-dimensional problem, in which the corresponding functions satisfy the uniform H\"{o}lder condition (see \cite{Strongin2000,Sergeyev2013})
-$$
-	\left|g_j(y(x'))-g_j(y(x''))\right| \leq K_j \left|x'-x'' \right|^{1/N}, \; x', x''\in [0,1], \; 1\leq j \leq m+1,
-$$
+\begin{align*} 
+	\left|g_j(y(x'))-g_j(y(x''))\right| \leq K_j \left|x'-x'' \right|^{1/N}, \\  
+	x', x''\in [0,1], \; 1\leq j \leq m+1,
+\end{align*}
 where $N$ is the dimensionality of the initial multidimensional problem and the coefficients $K_j$ are related to the Lipschitz constants $L_j$ of the initial problem by the relations
@@
 	According to (\ref{D_sets}) and (\ref{one_dimensional_problem}), the reduced one-dimensional functions $g_j (y(x)), 1 \leq j \leq m+1,$ will also be defined and computable in the corresponding intervals
-\begin{equation}\label{Q_intervals}
-	Q_1=[0,1], \; Q_{j+1}=\left\{x \in Q_j : g_j(y(x)) \leq 0 \right\}, \; 1 \leq j \leq m.
-\end{equation}
+\begin{align}\label{Q_intervals}
+	Q_1=[0,1], \; Q_{j+1}=\left\{x \in Q_j : g_j(y(x)) \leq 0 \right\}, \nonumber \\
+	1 \leq j \leq m.
+\end{align}
 These relations allow introducing a classification of the points $x \in [0,1]$ according to the number $\nu = \nu(x)$ of the constraints computable at the given point.
@@
 \end{equation}
-\begin{eqnarray} \nonumber
+\begin{eqnarray*}
 	\psi(x)=\frac{g_{\nu}(y(x))}{K_{\nu}} - 
@@
 	\right.
-\end{eqnarray}
+\end{eqnarray*}
 where $\nu = \nu(x)$ and $\varphi^\ast$ is the objective function value at the point of the problem solution (\ref{one_dimensional_problem}).
@@
 Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$, compute the \textit{characteristics} $R(i)$:
-\begin{equation}\label{R_1}
-	R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
-\end{equation}
+\begin{align}\label{R_1}
+	R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \nonumber \\ 
+ \nu=\nu(x_i)=\nu(x_{i-1}),
+\end{align}
 \begin{equation}\label{R_2}
@@
 Otherwise, execute a trial at the point
-$$
-	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
-$$
+\begin{align*}
+	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \\
+	\nu=\nu(x_{t-1})=\nu(x_t).
+\end{align*}
 The inequality $\Delta_t \leq \epsilon$ can be taken as the termination condition, where $t$ is from (\ref{MaxR}) and $\epsilon>0$ is the predefined accuracy.
@@
 Then, any limit point $\bar{y}$ of the sequence of the trials $\{y^k\}=\big\{y\big(x^k\big)\big\}$ generated by the index algorithm for problem (\ref{problem}) belongs to the set $Y_{\epsilon}$ form (\ref{Y_epsilon}) and satisfies the conditions
-$$
-	\varphi(\bar y) = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \leq \varphi(y_{\epsilon}).
-$$
 \end{enumerate}
+\begin{eqnarray*}
+	\varphi(\bar y) =	\inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \\ \leq \varphi(y_{\epsilon}).
+\end{eqnarray*}
 \textbf{Note 1}.
@@
 From here, it will follow that for any value $r_M>1$ the following inequalities will hold
-\begin{eqnarray}
-	R(i) & = & \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{z_i+z_{i-1}-2z_M^*}{r_M\mu_M} \geq \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\mu_M\Delta_i}{r_M\mu_M} \nonumber \\
-	& \geq & \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\Delta_i}{r_M} = \Delta_i + \Delta_i\frac{(z_i-z_{i-1})^2/\Delta_i^2}{r_M^2\mu_M^2\Delta_i^2} - 2\frac{\Delta_i}{r_M}
-\nonumber \\
-	& \geq & \Delta_i + \frac{\Delta_i}{r_M^2} - 2\frac{\Delta_i}{r_M} = \Delta_i \left( 1-\frac{1}{r_M} \right)^2>0.
-	\nonumber
-\end{eqnarray}
+\begin{align}
+	R(i) & = \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{z_i+z_{i-1}-2z_M^*}{r_M\mu_M} \nonumber \\
+	& \geq  \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\mu_M\Delta_i}{r_M\mu_M} \nonumber \\
+  & \geq \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\Delta_i}{r_M} \nonumber \\
+	& = \Delta_i + \Delta_i\frac{(z_i-z_{i-1})^2/\Delta_i^2}{r_M^2\mu_M^2\Delta_i^2} - 2\frac{\Delta_i}{r_M} \nonumber \\
+	& \geq \Delta_i + \frac{\Delta_i}{r_M^2} - 2\frac{\Delta_i}{r_M} = \Delta_i \left( 1-\frac{1}{r_M} \right)^2>0.
+\end{align}
 And because $r_{M}^{glob} > r_{M}^{loc} > 1$, then
-\begin{equation}\label{R_glob_greater_R_loc}
-	R_{glob}(i)=\Delta_i\left(1-\frac{1}{r_M^{glob}} \right)^2 > \Delta_i\left(1-\frac{1}{r_M^{loc}} \right)^2 = R_{loc}(i).
-\end{equation}
+\begin{align}\label{R_glob_greater_R_loc}
+	R_{glob}(i) & = \Delta_i\left(1-\frac{1}{r_M^{glob}} \right)^2 \nonumber \\
+	& > \Delta_i\left(1-\frac{1}{r_M^{loc}} \right)^2 = R_{loc}(i).
+\end{align}
 The statement is proved.
@@
 Otherwise, execute a trial at the point
-$$
-	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
-$$
+\begin{eqnarray*}
+	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \\
+	\nu=\nu(x_{t-1})=\nu(x_t).
+\end{eqnarray*}
 When computing the point $x^{k+1}$, use the value of the parameter $r_{\nu}$ fixed at Step~6.
@@
 	Let us illustrate the work of the algorithms IA and IA-DL when minimizing the function
-\begin{eqnarray}
-	\varphi(y_1, y_2)=-1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}- \nonumber \\
-	-\left(0.5(y_1-1)(y_2-1)\right)^4\exp{\left\{2-\left(0.5(y_1-1)\right)^4-(y_2-1)^4\right\}}
-	\nonumber
-\end{eqnarray}
+\begin{align*}
+	\varphi(y_1, y_2)&=-1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}-\\
+	&-\left(0.5(y_1-1)(y_2-1)\right)^4 \\
+	& \exp{\left\{2-\left(0.5(y_1-1)\right)^4-(y_2-1)^4\right\}}
+\end{align*}
 in the range $0 \leq y_1 \leq 4$, $1 \leq y_2 \leq 3$ subject to the constraints
@@
 satisfies the condition
-\begin{equation}\label{phi_bar_y}
-	\varphi(\bar y) = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \leq \varphi(y_{\epsilon}).
-\end{equation}
+\begin{align}\label{phi_bar_y}
+	\varphi(\bar y) & = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \nonumber \\
+	& \leq \varphi(y_{\epsilon}).
+\end{align}
 
@@
 	Let us show that if the conditions of Theorem 1 are satisfied, the inequality $R_{glob}(j)>0$ will hold. According to the second condition of Theorem 1 and to the condition (\ref{K_leq_L}), 
-$$
-	z_j=g_{\nu}\left( y(x_j) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{1/N}, \; \nu=\nu(x_j),
-$$
-$$
-	z_{j-1}=g_{\nu}\left( y(x_{j-1}) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \; \nu=\nu(x_{j-1}).
-$$
+\begin{align}
+	z_j &= g_{\nu}\left( y(x_j) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{1/N}, \nonumber \\
+	&\nu=\nu(x_j),
+\end{align}
+\begin{align}
+	z_{j-1} &= g_{\nu}\left( y(x_{j-1}) \right) \nonumber \\
+	&\leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \nonumber \\ 
+	&\nu=\nu(x_{j-1}).
+\end{align}
 
@@
 Indeed, at $\nu=\nu(x_j)=\nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
-\begin{eqnarray}
-	R(j) &=& \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^*}{r_{\nu}\mu_{\nu}}  \nonumber \\
-	&\geq& \Delta_i + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	&=& \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	&\geq& \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	&=& \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
-	\nonumber
-\end{eqnarray}
+\begin{align*}
+	R(j) &= \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \\
+	&\geq \Delta_i + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \\
+	&= \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^\ast-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
+	&\geq \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
+	&= \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
+\end{align*}
 At $\nu=\nu(x_j ) > \nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
-\begin{eqnarray}
-	R(j) &=& 2\Delta_i - 4\frac{z_j-z_{\nu}^*}{r_{\nu}\mu_{\nu}} \geq 2\Delta_i+4\frac{z_{\nu}^*-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	&=& 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
-	&\geq& 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
-	\nonumber
-\end{eqnarray}
+\begin{align*}
+	R(j) &= 2\Delta_i - 4\frac{z_j-z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \\
+	&\geq 2\Delta_i+4\frac{z_{\nu}^{\ast}-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} \\
+	&= 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
+	&\geq 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
+\end{align*}
 
@@
 \begin{table}[h!]
-	\centering
+	\small
+	%\centering
 	\caption{Average number of iterations performed by the methods on “Simple” class}
+	\label{tab:average_number_of_iterations_simple}
+
+		%\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}ccccc}
 		\begin{tabular}{ccccc}
@@
 			IA-DL, $r_{\nu}^{loc}=1.4$ & 316 & 7403 & 40419 & 296357 \\
-			\hline
+			%\hline
 			IA-DL, $r_{\nu}^{loc}=2.0$ & 350 & 8990(1) & 41230 & 360497 \\
-			\hline
+			%\hline
 			IA & 439 & 11804(1) & 41973(3) & 404660(2) \\
@@
 		\end{tabular}
-	\label{tab:average_number_of_iterations_simple}
+	\normalsize
 \end{table}
@@
 \begin{table}[h!]
-	\centering
+	\small
+	%\centering
 	\caption{Average number of iterations performed by the methods on “Hard” class}
+	\label{tab:average_number_of_iterations_hard}
+
 		\begin{tabular}{ccccc}
@@
 			IA-DL, $r_{\nu}^{loc}=1.4$ & 766 & 9636 & 60904 & 763711 \\
-			\hline
+			%\hline
 			IA-DL, $r_{\nu}^{loc}=2.0$ & 830 & 12020 & 57473 & 743874(1) \\
-			\hline
+			%\hline
 			IA & 1074 & 20327 & 87310(2) & 1084990(2) \\
@@
 		\end{tabular}
-	\label{tab:average_number_of_iterations_hard}
 \end{table}
@@
 
-% For one-column wide figures use
-\begin{figure}
-% Use the relevant command to insert your figure file.
-% For example, with the graphicx package use
-  \includegraphics{example.eps}
-% figure caption is below the figure
-\caption{Please write your figure caption here}
-\label{fig:1}       % Give a unique label
-\end{figure}
-%
-% For two-column wide figures use
-\begin{figure*}
-% Use the relevant command to insert your figure file.
-% For example, with the graphicx package use
-  \includegraphics[width=0.75\textwidth]{example.eps}
-% figure caption is below the figure
-\caption{Please write your figure caption here}
-\label{fig:2}       % Give a unique label
-\end{figure*}
-%
 % For tables use

### 2020-02-01 2019/NUMTA_Barkalov/SoftComputing/article.tex
Add address, fix tables and theorems.

@@
 %
-\usepackage{graphicx}
-\usepackage{lineno,hyperref,amsmath}
-%
 % \usepackage{mathptmx}      % use Times fonts if available on your TeX system
@@
 %\usepackage{latexsym}
+\usepackage{graphicx}
+\usepackage{lineno,hyperref,amsmath,array}
+\usepackage{multirow}
+\usepackage[round]{natbib} 
+\usepackage[misc]{ifsym}
 % etc.
@@
 
-\title{Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints%\thanks{Grants or other notes
-%about the article that should go on the front page should be
-%placed here. General acknowledgments should be placed at the end of the article.}
+\title{Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints\thanks{This research was supported by the Russian Science Foundation, project No.
+16-11-10150.}
 }
-%\subtitle{Do you have a subtitle?\\ If so, write it here}
-
-%\titlerunning{Short form of title}        % if too long for running head
 
@@
 
-%\authorrunning{Short form of author list} % if too long for running head
-
-\institute{F. Author \at
-              first address \\
-              Tel.: +123-45-678910\\
-              Fax: +123-45-678910\\
-              \email{fauthor@example.com}           %  \\
-%             \emph{Present address:} of F. Author  %  if needed
-           \and
-           S. Author \at
-              second address
+\institute{Roman Strongin        \and
+					 Konstantin Barkalov (\Letter) \and
+					 Semen Bevzuk \at
+              Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, 603950, Russia \\
+              \email{konstantin.barkalov@itmm.unn.ru}
 }
@@
 \begin{abstract}
-%Insert your abstract here. Include keywords, PACS and mathematical
-%subject classification numbers as needed.
 This paper considers the constrained global optimization problems, in which the functions are of the “black box” type and satisfy the Lipschitz condition.
@@
 \keywords{Global optimization \and Multiextremal problems \and Non-convex constraints \and Lipschitz constant estimates \and Dimensionality reduction \and Numerical methods}
-% \PACS{PACS code1 \and PACS code2 \and more}
-% \subclass{MSC code1 \and MSC code2 \and more}
+\subclass{MSC code1 \and MSC code2 \and more}
 \end{abstract}
@@
 	The main part of the paper has the following structure.
-In Section 2, the index method of accounting for the constraints and the dimensionality reduction scheme for the constrained global optimization problems are described.
-In Section 3, a description of the computational rules of the index global search algorithm is given and its theoretical properties are formulated.
-In Section 4, the index algorithm with dual Lipschitz constant estimates is formulated and its convergence is proved.
-Section 5 presents the results of the numerical experiments carried out on several series of the multiextremal problems with non-convex constraints.
-Section 6 concludes the paper.
+In Section \ref{sec:2}, the index method of accounting for the constraints and the dimensionality reduction scheme for the constrained global optimization problems are described.
+In Section \ref{sec:3}, a description of the computational rules of the index global search algorithm is given and its theoretical properties are formulated.
+In Section \ref{sec:4}, the index algorithm with dual Lipschitz constant estimates is formulated and its convergence is proved.
+Section \ref{sec:5} presents the results of the numerical experiments carried out on several series of the multiextremal problems with non-convex constraints.
+Section \ref{sec:6} concludes the paper.
 
 \section{Index Method of Accounting for the Constraints and Dimensionality Reduction}
-\label{sec:1}
+\label{sec:2}
 	We consider the global optimization problems of the form
@@
 \begin{align}\label{epsilon_reserved_solution}
-	\varphi(y_{\epsilon}) &= \nonumber \\
-	&\min{\left\{\varphi(y):y \in D, g_j(y)\leq -{\epsilon}_j, 1 \leq j \leq m\right\}},
+	\varphi(y_{\epsilon}) & = \nonumber \\
+	& \min{\left\{\varphi(y):y \in D, g_j(y)\leq -{\epsilon}_j, 1 \leq j \leq m\right\}},
 \end{align}
@@
 \begin{align}\label{Y_epsilon}
-	Y_{\epsilon} &= \nonumber \\
-	&\left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
+	Y_{\epsilon} & = \nonumber \\
+	& \left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
 \end{align}
@@
 
-\begin{figure}%[h]
+\begin{figure}
 	\center{\includegraphics[width=0.6\linewidth]{epsilon_reserved_solution}}
@@
 \begin{align}\label{one_dimensional_problem}
-	\varphi(y(x^\ast)) =  
-	\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
+	& \varphi(y(x^\ast)) =  \nonumber \\
+	& \min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
 \end{align}
@@
 For  illustration, Fig.
-\ref{fig:evolvent} (a), (b) shows the projections of the interval $[0,1]$ for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
+\ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
 \begin{figure}
@@
 \section{Index-Based Global Search Algorithm}
-\label{sec:2}
+\label{sec:3}
 	The index-based global search algorithm under consideration for solving the problem (\ref{problem}) (here it is given according to \cite{Strongin2000}) involves the construction of a sequence of points $x^k \in (0,1)$, at which the search trials are conducted with the results of the type (\ref{trial_result}).
@@
 \\
-\textbf{Theorem 1}. Assume that the following conditions are satisfied:
+\begin{theorem}\label{theorem:1} Assume that the following conditions are satisfied:
 	\begin{enumerate} 
@@
 			\end{equation}
+	\end{enumerate}
 	Then, any limit point $\bar{y}$ of the sequence of the trials $\{y^k\}=\big\{y\big(x^k\big)\big\}$ generated by the index algorithm for problem (\ref{problem}) belongs to the set $Y_{\epsilon}$ form (\ref{Y_epsilon}) and satisfies the conditions
-\end{enumerate}
 	\begin{eqnarray*}
-	\varphi(\bar y) =	\inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \\ \leq \varphi(y_{\epsilon}).
+		\varphi(\bar y) =	\inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \\
+		\leq \varphi(y_{\epsilon}).
 	\end{eqnarray*}
+\end{theorem}
 \textbf{Note 1}.
@@
 \section{Index Algorithm with Dual Lipschitz Constant Estimates}
-\label{sec:3}
+\label{sec:4}
 	The index global search algorithm described in the previous section (hereafter referred to as IA) is intended for solving  the multiextremal problems, in which the objective function and the constraints satisfy the Lipschitz condition.
@@
 
-	According to Theorem 1, the sequence of trial points $\{y^k\}$ will converge to a point $\bar y \in Y_\epsilon$ if the condition (\ref{theorem_inequalities}) is satisfied.
+	According to Theorem \ref{theorem:1}, the sequence of trial points $\{y^k\}$ will converge to a point $\bar y \in Y_\epsilon$ if the condition (\ref{theorem_inequalities}) is satisfied.
 So far, an appropriate choice of the parameters $r_{\nu}$ allows using the values
@@
 
-	Note that Theorem 1 on the convergence of the index method implies the generation of an infinite sequence of trials $\{y^k\}=\big\{y(x^k )\big\}$ by the algorithm.
+	Note that Theorem \ref{theorem:1} on the convergence of the index method implies the generation of an infinite sequence of trials $\{y^k\}=\big\{y(x^k )\big\}$ by the algorithm.
 In this case, at the final stage of the search (i.e.
@@
 
-	The first four steps of the algorithm repeat Step 1 -- Step 4 of the initial index algorithm from Section 3.
+	The first four steps of the algorithm repeat Step 1 -- Step 4 of the initial index algorithm from Section \ref{sec:3}.
 
@@
 	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \\
-	\nu=\nu(x_{t-1})=\nu(x_t).
+	\nu = \nu(x_{t-1})=\nu(x_t).
 \end{eqnarray*}
@@
 \begin{align*}
-	\varphi(y_1, y_2)&=-1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}-\\
-	&-\left(0.5(y_1-1)(y_2-1)\right)^4 \\
+	\varphi(y_1, y_2) & = -1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}-\\
+	& -\left(0.5(y_1-1)(y_2-1)\right)^4 \\
 	& \exp{\left\{2-\left(0.5(y_1-1)\right)^4-(y_2-1)^4\right\}}
@@
 in the range $0 \leq y_1 \leq 4$, $1 \leq y_2 \leq 3$ subject to the constraints
-\begin{eqnarray}
-	g_1(y_1, y_2) &=& 0.01 \left( (y_1-2.2)^2+(y_2-1.2)^2-2.25 \right) \leq 0, \nonumber \\
-	g_2(y_1, y_2) &=& 100 \left(1-(y_1-2)^2/1.44+(0.5y_2)^2 \right) \leq 0, \nonumber \\
-	g_3(y_1, y_2) &=& 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0. \nonumber
-\end{eqnarray}
+\begin{eqnarray*}
+	g_1(y_1, y_2) & = & 0.01 \left( (y_1-2.2)^2+(y_2-1.2)^2-2.25 \right) \leq 0, \\
+	g_2(y_1, y_2) & = & 100 \left(1-(y_1-2)^2/1.44+(0.5y_2)^2 \right) \leq 0, \\
+	g_3(y_1, y_2) & = & 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0.
+\end{eqnarray*}
 The feasible domain of this problem is non-connected and consists of three subdomains.
-The global minimum $\varphi(y^\ast)=-1.489$ is achieved at the point $y^\ast =(0.942, 0.944)$ (the values are given with the accuracy of $10^{-3}$).
+The global minimum $\varphi(y^\ast)=-1.489$ is achieved at the point \\ $y^\ast =(0.942, 0.944)$ (the values are given with the accuracy of $10^{-3}$).
 
@@
 To solve the problem with a predefined accuracy $\epsilon=0.002$, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method.
-The results of larger-scale numerical experiments are presented in Section 4.
+The results of larger-scale numerical experiments are presented in Section \ref{sec:4}.
 
 	Let us formulate and prove the convergence conditions for the new algorithm IA-DL.\\
-\textbf{Theorem 2}.
-Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem 1 are fulfilled.
+\begin{theorem}\label{theorem:2} Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem \ref{theorem:1} are fulfilled.
 Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1,\; 1\leq \nu \leq m+1,$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e.
@@
 \end{align}
+\end{theorem}
 
 \textbf{Proof}.
-Because for the IA with the parameters $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem 1 are fulfilled, then any limit point $\bar{y}=y(\bar{x})$ of the trial sequence generated by the IA for the problem (\ref{problem}) satisfies the conditions (\ref{phi_bar_y}).
+Because for the IA with the parameters $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem \ref{theorem:1} are fulfilled, then any limit point $\bar{y}=y(\bar{x})$ of the trial sequence generated by the IA for the problem (\ref{problem}) satisfies the conditions (\ref{phi_bar_y}).
 
@@
 
-	Let us show that if the conditions of Theorem 1 are satisfied, the inequality $R_{glob}(j)>0$ will hold. According to the second condition of Theorem 1 and to the condition (\ref{K_leq_L}), 
+	Let us show that if the conditions of Theorem \ref{theorem:1} are satisfied, the inequality $R_{glob}(j)>0$ will hold.
+According to the second condition of Theorem \ref{theorem:1} and to the condition (\ref{K_leq_L}),
 \begin{align}
-	z_j &= g_{\nu}\left( y(x_j) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{1/N}, \nonumber \\
-	&\nu=\nu(x_j),
+	z_j & = g_{\nu}\left( y(x_j) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{1/N}, \nonumber \\
+	& \nu=\nu(x_j),
 \end{align}
@@
 	z_{j-1} &= g_{\nu}\left( y(x_{j-1}) \right) \nonumber \\
-	&\leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \nonumber \\ 
-	&\nu=\nu(x_{j-1}).
+	& \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \nonumber \\ 
+	& \nu=\nu(x_{j-1}).
 \end{align}
@@
 \begin{align*}
-	R(j) &= \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \\
-	&\geq \Delta_i + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \\
-	&= \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^\ast-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
-	&\geq \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
-	&= \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
+	& R(j) = \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \\
+	& \geq \Delta_i + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \\
+	& = \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^\ast-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
+	& \geq \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
+	& = \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
 \end{align*}
@@
 \begin{align*}
-	R(j) &= 2\Delta_i - 4\frac{z_j-z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \\
-	&\geq 2\Delta_i+4\frac{z_{\nu}^{\ast}-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} \\
-	&= 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
-	&\geq 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
+	& R(j) = 2\Delta_i - 4\frac{z_j-z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \\
+	& \geq 2\Delta_i+4\frac{z_{\nu}^{\ast}-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} \\
+	& = 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
+	& \geq 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
 \end{align*}
@@
 \section{Numerical experiments}
-\label{sec:4}
+\label{sec:5}
 To demonstrate the efficiency of the proposed algorithm, let us consider the results of solving several series of multidimensional constrained global optimization problems.
@@
 \begin{table}[h!]
-	\small
-	%\centering
 	\caption{Average number of iterations performed by the methods on “Simple” class}
 	\label{tab:average_number_of_iterations_simple}
-
-		%\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}ccccc}
-		\begin{tabular}{ccccc}
-			\hline
+	\begin{tabular}[]{ccccc}
+		\hline\noalign{\smallskip}
 		                    & $N=2$                & $N=3$                    & $N=4$                  & $N=5$                   \\
-			\hline
-			IA-DL, $r_{\nu}^{loc}=1.4$ & 316 & 7403 & 40419 & 296357 \\
+		\noalign{\smallskip}\hline\noalign{\smallskip}
+		IA-DL,              & \multirow{2}{*}{316} & \multirow{2}{*}{7403}    & \multirow{2}{*}{40419} & \multirow{2}{*}{296357} \\
+		$r_{\nu}^{loc}=1.4$ &                      &                          &                        &                         \\
 		%\hline
-			IA-DL, $r_{\nu}^{loc}=2.0$ & 350 & 8990(1) & 41230 & 360497 \\
+		IA-DL,              & \multirow{2}{*}{350} & \multirow{2}{*}{8990(1)} & \multirow{2}{*}{41230} & \multirow{2}{*}{360497} \\
+		$r_{\nu}^{loc}=2.0$ &                      &                          &                        &                         \\
 		%\hline
 		IA                  & 439                  & 11804(1)                 & 41973(3)               & 404660(2)               \\
-			\hline
+		\hline\noalign{\smallskip}
 	\end{tabular}
-	\normalsize
 \end{table}
@@
 \begin{table}[h!]
-	\small
-	%\centering
 	\caption{Average number of iterations performed by the methods on “Hard” class}
 	\label{tab:average_number_of_iterations_hard}
-
 	\begin{tabular}{ccccc}
-			\hline
+		\hline\noalign{\smallskip}
 		                    & $N=2$                & $N=3$                    & $N=4$                  & $N=5$                      \\
-			\hline
-			IA-DL, $r_{\nu}^{loc}=1.4$ & 766 & 9636 & 60904 & 763711 \\
+		\noalign{\smallskip}\hline\noalign{\smallskip}
+		IA-DL,              & \multirow{2}{*}{766} & \multirow{2}{*}{9636}    & \multirow{2}{*}{60904} & \multirow{2}{*}{763711}    \\
+		$r_{\nu}^{loc}=1.4$ &                      &                          &                        &                            \\
 		%\hline
-			IA-DL, $r_{\nu}^{loc}=2.0$ & 830 & 12020 & 57473 & 743874(1) \\
+		IA-DL,              & \multirow{2}{*}{830} & \multirow{2}{*}{12020}   & \multirow{2}{*}{57473} & \multirow{2}{*}{743874(1)} \\
+		$r_{\nu}^{loc}=2.0$ &                      &                          &                        &                            \\
 		%\hline
 		IA                  & 1074                 & 20327                    & 87310(2)               & 1084990(2)                 \\
-			\hline
+		\hline\noalign{\smallskip}
 	\end{tabular}
@@
 \section{Conclusion}
-\label{sec:5}
+\label{sec:6}
 	In this paper, a method for solving the Lipschitz optimization problems, which constitute an important class of the global optimization problems, was considered.
@@
 
-\section{Section title}
-\label{sec:10}
-Text with citations \cite{RefB} and \cite{RefJ}.
-\subsection{Subsection title}
-\label{sec:20}
-as required. Don't forget to give each section
-and subsection a unique label (see Sect.~\ref{sec:1}).
-\paragraph{Paragraph headings} Use paragraph headings as needed.
-\begin{equation}
-a^2+b^2=c^2
-\end{equation}
-
-% For tables use
-\begin{table}
-% table caption is above the table
-\caption{Please write your table caption here}
-\label{tab:1}       % Give a unique label
-% For LaTeX tables use
-\begin{tabular}{lll}
-\hline\noalign{\smallskip}
-first & second & third  \\
-\noalign{\smallskip}\hline\noalign{\smallskip}
-number & number & number \\
-number & number & number \\
-\noalign{\smallskip}\hline
-\end{tabular}
-\end{table}
-
 \begin{acknowledgements}
@@
 
-% Authors must disclose all relationships or interests that 
-% could have direct or potential influence or impart bias on 
-% the work: 
-%
-% \section*{Conflict of interest}
-%
-% The authors declare that they have no conflict of interest.
-
 % BibTeX users please use one of
 %\bibliographystyle{spbasic}      % basic style, author-year citations
-%\bibliographystyle{spmpsci}      % mathematics and physical sciences
+%\bibliographystyle{spbasic}      % mathematics and physical sciences
 %\bibliographystyle{spphys}       % APS-like style for physics
 %\bibliography{}   % name your BibTeX data base
-\bibliographystyle{spmpsci}
+\bibliographystyle{spbasic}
 \bibliography{bibliography}

### 2020-02-01 2019/NUMTA_Barkalov/SoftComputing/article.tex
Decreased the font in 4, 5

@@
 	Along with the exact solution $y^\ast$ of the problem (\ref{problem}), we will consider also the $\varepsilon$-reserved solution $y_{\epsilon}$, for which 
+\small
 \begin{align}\label{epsilon_reserved_solution}
-	\varphi(y_{\epsilon}) & = \nonumber \\
-	& \min{\left\{\varphi(y):y \in D, g_j(y)\leq -{\epsilon}_j, 1 \leq j \leq m\right\}},
+	\varphi(y_{\epsilon}) = \min{\left\{\varphi(y):y \in D, g_j(y)\leq -{\epsilon}_j, 1 \leq j \leq m\right\}},
 \end{align}
+\normalsize
 and the set of solutions $Y_{\epsilon}$, which are not worse than the $\varepsilon$-reserved solution in the objective function value, i.e.
+\small
 \begin{align}\label{Y_epsilon}
-	Y_{\epsilon} & = \nonumber \\
-	& \left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
+	Y_{\epsilon} = \left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
 \end{align}
+\normalsize
 Fig. \ref{fig:eps_reserved_solution} illustrates the concepts introduced.

### 2020-02-02 2019/NUMTA_Barkalov/SoftComputing/article.tex
Renamed illustrations: Name your figure files with Fig and the figure number
correct some references
Changed some formulas

@@
 \RequirePackage{fix-cm}
-%
-%\documentclass{svjour3}                     % onecolumn (standard format)
-%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
-%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
+
 \documentclass[twocolumn]{svjour3}          % twocolumn
-%
+
 \smartqed  % flush right qed marks, e.g. at end of proof
@@
 \usepackage{graphicx}
-\usepackage{lineno,hyperref,amsmath,array}
+\usepackage{lineno,hyperref,amsmath}
 \usepackage{multirow}
@@
 \keywords{Global optimization \and Multiextremal problems \and Non-convex constraints \and Lipschitz constant estimates \and Dimensionality reduction \and Numerical methods}
-\subclass{MSC code1 \and MSC code2 \and more}
+%\subclass{90C26}
 \end{abstract}
@@
 
-	It is supposed that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_{j}(y), 1 \leq j \leq m$, of the constraints satisfy the Lipschitz condition
+	It is supposed that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_{j}(y)$, $1 \leq j \leq m$, of the constraints satisfy the Lipschitz condition
 \begin{align}\label{lipschitz_condition}
@@
 \begin{figure}
-	\center{\includegraphics[width=0.6\linewidth]{epsilon_reserved_solution}}
+	\center{\includegraphics[width=0.6\linewidth]{Fig1}}
 	\caption{$\varepsilon$-reserved solution.}
@@
 
-	We will consider the initial problem (\ref{problem}) assuming that the problem functions $g_{j}(y), 1 \leq j \leq m+1,$ may be defined only partly.
+	We will consider the initial problem (\ref{problem}) assuming that the problem functions $g_{j}(y)$, $1 \leq j \leq m+1$, may be defined only partly.
 It means that they are defined and computable only in the sets
+\small
 \begin{align}\label{D_sets}
-	D_1 = D, \; D_{j+1} = \left\{ y \in D_j: \; g_j(y) \leq 0 \right\}, \nonumber \\
-	1 \leq j \leq m.
+	D_1 = D, \; D_{j+1} = \left\{ y \in D_j: \; g_j(y) \leq 0 \right\}, \; 1 \leq j \leq m.
 \end{align}
+\normalsize
 This situation is typical for the optimal design problems when some conditions for the functioning of the modeled object are not met and its other characteristics appear to be undefined.
@@
 Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
+\small
 \begin{align}\label{one_dimensional_problem}
@@
 \end{align}
+\normalsize
 The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000,Sergeyev2013}.
@@
 	\begin{minipage}[h]{0.45\linewidth}
-		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=4} \\ {\footnotesize(a)}}
+		\center{\includegraphics[width=1.0\linewidth]{Fig2} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.45\linewidth}
-		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=5} \\ {\footnotesize(b)}}
+		\center{\includegraphics[width=1.0\linewidth]{Fig3} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 	\begin{eqnarray*}
-		\varphi(\bar y) =	\inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \\
-		\leq \varphi(y_{\epsilon}).
+		\varphi(\bar y) & = &	\inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \\
+		& \leq & \varphi(y_{\epsilon}).
 	\end{eqnarray*}
@@
 	Let us illustrate the work of the algorithms IA and IA-DL when minimizing the function
+\small
 \begin{align*}
-	\varphi(y_1, y_2) & = -1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}-\\
-	& -\left(0.5(y_1-1)(y_2-1)\right)^4 \\
-	& \exp{\left\{2-\left(0.5(y_1-1)\right)^4-(y_2-1)^4\right\}}
+	&\varphi(y_1, y_2) = -1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}-\\
+	& -\left(0.5(y_1-1)(y_2-1)\right)^4 \exp{\left\{2-\left(0.5(y_1-1)\right)^4-(y_2-1)^4\right\}}
 \end{align*}
+\normalsize
 in the range $0 \leq y_1 \leq 4$, $1 \leq y_2 \leq 3$ subject to the constraints
+\small
 \begin{eqnarray*}
@@
 \end{eqnarray*}
+\normalsize
 The feasible domain of this problem is non-connected and consists of three subdomains.
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{stronginC3_IA} \\ {\footnotesize(a)}}
+		\center{\includegraphics[width=\linewidth]{Fig4} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{stronginC3_IA-DL} \\ {\footnotesize(b)}}
+		\center{\includegraphics[width=\linewidth]{Fig5} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 \begin{align}
-	z_j & = g_{\nu}\left( y(x_j) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{1/N}, \nonumber \\
-	& \nu=\nu(x_j),
+	& z_j = g_{\nu}\left( y(x_j) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{1/N}, \nonumber \\
+	& \nu = \nu(x_j),
 \end{align}
 \begin{align}
-	z_{j-1} &= g_{\nu}\left( y(x_{j-1}) \right) \nonumber \\
+	z_{j-1} & = g_{\nu}\left( y(x_{j-1}) \right) \nonumber \\
 	& \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \nonumber \\
-	& \nu=\nu(x_{j-1}).
+	& \nu = \nu(x_{j-1}).
 \end{align}
@@
 	Since the point $\bar y = y(\bar x)$ belongs to the set $Y_{\epsilon}$, then
-$$
+\begin{equation*}
 	g_{\nu}\left( y(\bar x) \right) \leq -\epsilon_{\nu}, \; 1\leq\nu\leq m,
-$$
-$$
+\end{equation*}
+\begin{equation*}
 	g_{m+1}\left( y(\bar x) \right) \leq z_{m+1}^\ast.
-$$
+\end{equation*}
 
-	Then, the characteristic $R_{glob}(j)$ of the interval $[x_{j-1}, x_j], j=j(k)$ will be strictly positive, i.e.
-$R_{glob}(j)>0$.
-Indeed, at $\nu=\nu(x_j)=\nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
+	Then, the characteristic $R_{glob}(j)$ of the interval \\
+$[x_{j-1}, x_j]$, $j=j(k)$ will be strictly positive, i.e. \\
+$R_{glob}(j)>0$. Indeed, at $\nu=\nu(x_j)=\nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
+\small
 \begin{align*}
-	& R(j) = \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \\
-	& \geq \Delta_i + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \\
-	& = \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^\ast-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
-	& \geq \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
+	& R(j) = \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \geq \Delta_i + \\ 
+	& + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \\
+	& = \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} + \\
+	& + 4\frac{z_{\nu}^\ast-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
+	& \geq \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right) + \\
+	& + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
 	& = \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
 \end{align*}
+\normalsize
 At $\nu=\nu(x_j ) > \nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Simple_59_IA} \\ {\footnotesize(a)}}
+		\center{\includegraphics[width=\linewidth]{Fig6} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Simple_59_IA-DL} \\ {\footnotesize(b)}}
+		\center{\includegraphics[width=\linewidth]{Fig7} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Hard_7_IA} \\ {\footnotesize(a)}}
+		\center{\includegraphics[width=\linewidth]{Fig8} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Hard_7_IA-DL} \\ {\footnotesize(b)}}
+		\center{\includegraphics[width=\linewidth]{Fig9} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_3} \\ {\footnotesize(a)}}
+		\center{\includegraphics[width=\linewidth]{Fig10} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_3} \\ {\footnotesize(b)}}
+		\center{\includegraphics[width=\linewidth]{Fig11} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_5} \\ {\footnotesize(a)}}
+		\center{\includegraphics[width=\linewidth]{Fig12} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_5} \\ {\footnotesize(b)}}
+		\center{\includegraphics[width=\linewidth]{Fig13} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Decision_rule_gklssC_4} \\ {\footnotesize(a)}}
+		\center{\includegraphics[width=\linewidth]{Fig14} \\ {\footnotesize(a)}}
 	\end{minipage}
@@
 	\begin{minipage}[h]{0.48\linewidth}
-		\center{\includegraphics[width=\linewidth]{Decision_rule_gklshC_4} \\ {\footnotesize(b)}}
+		\center{\includegraphics[width=\linewidth]{Fig15} \\ {\footnotesize(b)}}
 	\end{minipage}
@@
 %\bibliographystyle{spphys}       % APS-like style for physics
-%\bibliography{}   % name your BibTeX data base
 \bibliographystyle{spbasic}

### 2020-02-03 2019/NUMTA_Barkalov/SoftComputing/article.tex
add compliance section and correct references

@@
 In this paper, we consider constrained global optimization problems and the numerical methods for solving problems of this type.
-Such problems are often encountered in applications (see, for example, the review \cite{Pinter2006}).
+Such problems are often encountered in applications (see, for example, \cite{Pinter2006}).
 As a rule, problem functions are defined by a program code, i.e.
@@
 	The methods for solving the Lipschitzian optimization problems constitute a very important area in the development of the global optimization methods and are the objects of investigations of many researchers.
-The first algorithms of this class were proposed as early as in the 1970s \cite{Evtushenko1971,Piyavskii1972,Shubert1972,Strongin1970}; since that time, this area has been continuing to develop extensively (see, for example \cite{Evtushenko2009,Evtushenko2013,Strongin2000,Sergeyev2013,Jones2009}).
-Note that these global optimization algorithms are, as a rule, oriented towards solving either  unconstrained optimization problems
-	\cite{Paulavicius2014,Sergeyev2015,Pinter1996,Jones1993,Gablonsky2001}
-or problems with constraints of a simple structure \cite{Vaz2009,Paulavicius2016}.
-The solving of the problems with complex non-convex constraints is usually performed with the use of the penalty functions method \cite{Stripinis2019,Pillo2012,Pillo2016}, which has a number of drawbacks.
+The first algorithms of this class were proposed as early as in the 1970s (\cite{Evtushenko1971,Piyavskii1972,Shubert1972,Strongin1970}); since that time, this area has been continuing to develop extensively (see, for example, \cite{Evtushenko2009,Evtushenko2013,Strongin2000,Sergeyev2013,Jones2009}).
+Note that these global optimization algorithms are, as a rule, oriented towards solving either  unconstrained optimization problems \break (\cite{Paulavicius2014,Sergeyev2015,Pinter1996,Jones1993,Gablonsky2001}) or problems with constraints of a simple structure (\cite{Vaz2009,Paulavicius2016}).
+The solving of the problems with complex non-convex constraints is usually performed with the use of the penalty functions method (\cite{Stripinis2019,Pillo2012,Pillo2016}), which has a number of drawbacks.
 In particular, this method cannot be applied to the problems with the partially defined functions (i.e.
@@
 	Within the framework of our research, an original approach to the minimization of the multiextremal functions with non-convex constraints called the index method of the accounting for the constraints
-	\cite{Strongin2000,Pugliese,Barkalov2002,Strongin2003} 
-was used.
+was used (see \cite{Strongin2000,Pugliese,Barkalov2002,Strongin2003}).
 This approach is based on separate accounting for each constraint of the problem and does not involve the use of the penalty functions.
@@
 This allows solving the problems with partially defined functions.
-The experimental comparison of the index method of accounting for the constraints and the penalty function method demonstrated the advantages of the index method in terms of the number of iterations and the number of computations of the problem functions \cite{Barkalov2017_1,Barkalov2017_2}.
+The experimental comparison of the index method of accounting for the constraints and the penalty function method demonstrated the advantages of the index method in terms of the number of iterations and the number of computations of the problem functions (\cite{Barkalov2017_1,Barkalov2017_2}).
 
@@
 \begin{itemize}
-\item global estimate of the constant $L$ in the whole search domain $D$ \cite{Strongin2000,Pinter1996,Horst1996}.
-\item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ \cite{Kvasov2003,Sergeyev2010,Sergeyev2016}.
-\item choice of the estimates of the constant $L$ from some set of possible values \cite{Jones2009,Jones1993,Gablonsky2001,Sergeyev2006}.
+\item global estimate of the constant $L$ in the whole search domain $D$ (\cite{Strongin2000,Pinter1996,Horst1996}).
+\item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ (\cite{Kvasov2003,Sergeyev2010,Sergeyev2016}).
+\item choice of the estimates of the constant $L$ from some set of possible values (\cite{Jones2009,Jones1993,Gablonsky2001,Sergeyev2006}).
 \end{itemize}
@@
 The choice of one of the two estimates to be used in the rules of the algorithm is performed in the adaptive manner, depending on the behavior of the function.
-This work continues the research reported in the paper
-	\cite{NUMTA2019},
-in which the preliminary results have been obtained for the unconstrained problems.
+This work continues the research reported in \cite{NUMTA2019}, in which the preliminary results have been obtained for the unconstrained problems.
 
@@
 
-	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions \cite{Sergeyev2006} or simplicial partitions \cite{Zilinskas2008}).
+	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions in \cite{Sergeyev2006} or simplicial partitions in \cite{Zilinskas2008}).
 In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve).
@@
 	Since we have assumed that $x'$ is a limit point for the IM-DL and  $x' \neq \bar x$, then there exists such a sequence of the intervals $[x_{t-1}, x_t], t=t(k)$ that $x' \in [x_{t-1}, x_t], R_{loc}(t)>0$, and $R_{loc}(t) \rightarrow 0$.
-This property of the characteristics of the intervals shrinking to the limit point has been established when proving the theorem on the convergence of the IA in \cite{Sergeyev2013}.
+This property of the characteristics of the intervals shrinking to the limit point has been established when proving the theorem on the convergence of the IA in \cite{Strongin2000}.
 
@@
 To demonstrate the efficiency of the proposed algorithm, let us consider the results of solving several series of multidimensional constrained global optimization problems.
-In order to construct the problems, the functions generated by the GKLS generator \cite{Gaviano2003} were used.
-This generator of multiextremal functions is often used for investigating global optimization algorithms \cite{Paulavicius2014,Sergeyev2015,Barkalov2018}.
+In order to construct the problems, the functions generated by the GKLS generator (\cite{Gaviano2003}) were used.
+This generator of multiextremal functions is often used for investigating global optimization algorithms (\cite{Sergeyev2015,Paulavicius2014,Barkalov2018}).
 According to \cite{Gaviano2003}, we used simpler and more complex functions, which are  further denoted as ``Simple'' and ``Hard'', respectively.
@@
 
-	Further research areas include the use of the dual estimate of the Lipschitz constant in parallel algorithms, which can be constructed with the use of various schemes for  dimensionality reduction and  for parallelization of the computations \cite{Strongin2018}.
+	Further research areas include the use of the dual estimate of the Lipschitz constant in parallel algorithms, which can be constructed with the use of various schemes for  dimensionality reduction and  for parallelization of the computations (\cite{Strongin2018}).
 
-\begin{acknowledgements}
-This research was supported by the Russian Science Foundation, project No. 16-11-10150.
-\end{acknowledgements}
+%\begin{acknowledgements}
+%This research was supported by the Russian Science Foundation, project No. 16-11-10150.
+%\end{acknowledgements}
+
+\section*{Compliance with Ethical Standards}
+\textbf{Conflict of interest} The authors declare that they have no conflict of interest.
+
+\textbf{Ethical approval} This article does not contain any studies with human participants or
+animals performed by any of the authors.
 

### 2020-02-03 2019/NUMTA_Barkalov/SoftComputing/article.tex
fix formulas

@@
 \begin{align}
-	R(i) & = \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{z_i+z_{i-1}-2z_M^*}{r_M\mu_M} \nonumber \\
-	& \geq  \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\mu_M\Delta_i}{r_M\mu_M} \nonumber \\
-  & \geq \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\Delta_i}{r_M} \nonumber \\
-	& = \Delta_i + \Delta_i\frac{(z_i-z_{i-1})^2/\Delta_i^2}{r_M^2\mu_M^2\Delta_i^2} - 2\frac{\Delta_i}{r_M} \nonumber \\
+	R(i) & = \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{z_i+z_{i-1}-2z_M^*}{r_M\mu_M} \geq \nonumber \\
+	& \geq  \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\mu_M\Delta_i}{r_M\mu_M} \geq \nonumber \\
+  & \geq \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\Delta_i}{r_M} = \nonumber \\
+	& = \Delta_i + \Delta_i\frac{(z_i-z_{i-1})^2/\Delta_i^2}{r_M^2\mu_M^2\Delta_i^2} - 2\frac{\Delta_i}{r_M} \geq \nonumber \\
 	& \geq \Delta_i + \frac{\Delta_i}{r_M^2} - 2\frac{\Delta_i}{r_M} = \Delta_i \left( 1-\frac{1}{r_M} \right)^2>0.
@@
 \begin{align}\label{R_glob_greater_R_loc}
-	R_{glob}(i) & = \Delta_i\left(1-\frac{1}{r_M^{glob}} \right)^2 \nonumber \\
+	R_{glob}(i) & = \Delta_i\left(1-\frac{1}{r_M^{glob}} \right)^2 > \nonumber \\
 	& > \Delta_i\left(1-\frac{1}{r_M^{loc}} \right)^2 = R_{loc}(i).
@@
 \begin{align}
-	z_{j-1} & = g_{\nu}\left( y(x_{j-1}) \right) \nonumber \\
+	z_{j-1} & = g_{\nu}\left( y(x_{j-1}) \right) \leq \nonumber \\
 	& \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \nonumber \\
@@
 
-	Then, the characteristic $R_{glob}(j)$ of the interval \\
-$[x_{j-1}, x_j]$, $j=j(k)$ will be strictly positive, i.e. \\
+	Then, the characteristic $R_{glob}(j)$ of the interval \break
+$[x_{j-1}, x_j]$, $j=j(k)$ will be strictly positive, i.e. \break
 $R_{glob}(j)>0$. Indeed, at $\nu=\nu(x_j)=\nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
@@
 	& R(j) = \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \geq \Delta_i + \\ 
-	& + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \\
+	& + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}}\\
 	& = \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} + \\
-	& + 4\frac{z_{\nu}^\ast-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
+	& + 4\frac{z_{\nu}^\ast-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \geq \\
 	& \geq \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right) + \\
-	& + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
+	& + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} = \\
 	& = \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
@@
 \begin{align*}
-	& R(j) = 2\Delta_i - 4\frac{z_j-z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \\
-	& \geq 2\Delta_i+4\frac{z_{\nu}^{\ast}-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} \\
-	& = 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
+	& R(j) = 2\Delta_i - 4\frac{z_j-z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \geq \\
+	& \geq 2\Delta_i+4\frac{z_{\nu}^{\ast}-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} = \\
+	& = 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \geq \\
 	& \geq 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
@@
 		$r_{\nu}^{loc}=1.4$ &                      &                          &                        &                         \\
-		%\hline
+		\noalign{\smallskip}
 		IA-DL,              & \multirow{2}{*}{350} & \multirow{2}{*}{8990(1)} & \multirow{2}{*}{41230} & \multirow{2}{*}{360497} \\
 		$r_{\nu}^{loc}=2.0$ &                      &                          &                        &                         \\
-		%\hline
+		\noalign{\smallskip}
 		IA                  & 439                  & 11804(1)                 & 41973(3)               & 404660(2)               \\
@@
 		$r_{\nu}^{loc}=1.4$ &                      &                          &                        &                            \\
-		%\hline
+		\noalign{\smallskip}
 		IA-DL,              & \multirow{2}{*}{830} & \multirow{2}{*}{12020}   & \multirow{2}{*}{57473} & \multirow{2}{*}{743874(1)} \\
 		$r_{\nu}^{loc}=2.0$ &                      &                          &                        &                            \\
-		%\hline
+		\noalign{\smallskip}
 		IA                  & 1074                 & 20327                    & 87310(2)               & 1084990(2)                 \\

### 2020-02-04 2019/NUMTA_Barkalov/SoftComputing/article.tex
last correction before submission

@@
 	\end{enumerate}
-	Then, any limit point $\bar{y}$ of the sequence of the trials $\{y^k\}=\big\{y\big(x^k\big)\big\}$ generated by the index algorithm for problem (\ref{problem}) belongs to the set $Y_{\epsilon}$ form (\ref{Y_epsilon}) and satisfies the conditions
+	Then, any limit point $\bar{y}$ of the sequence of the trials $\{y^k\}=\big\{y\big(x^k\big)\big\}$ generated by the index algorithm for problem (\ref{problem}) belongs to the set $Y_{\epsilon}$ from (\ref{Y_epsilon}) and satisfies the conditions
 	\begin{eqnarray*}

### 2020-02-04 2020/PAVT_Poster/poster-ru.tex
poster final version

@@
 С помощью параллельного АГП-Р проведена серия экспериментов на тестовых задачах с разрывными функциями.
-Было получено линейное сокращение числа итераций метода в зависимости от числа задействованных потоков, что в предположении большой трудоемкости вычислений значений целевой функции будет соответствовать линейному ускорению по времени.
-Например, при использовании 64 потоков (на одном узле кластера) ускорение по итерациям составляло от 50 до 55 в зависимости от решаемой задачи.
+Наблюдалось линейное сокращение числа итераций метода в зависимости от числа задействованных потоков, что в предположении большой трудоемкости вычислений значений целевой функции будет соответствовать линейному ускорению по времени.
+Например, при использовании (на одном узле кластера) 32 потоков было получено ускорение по итерациям в 25 раз, а при использовании 64 потоков --- в 50 раз.
 

### 2020-02-06 2020/PAVT_Sovrasov/pavt_paper.tex
PAVT: fix bibliorgaphy style
PAVT: fix style again

@@
 \usepackage{cite}
+\usepackage{doi}
 \usepackage{subfig}
@@
 \newtheorem{theorem}{Теорема}
+\renewcommand{\doitext}{DOI: }
 
@@
     {\includegraphics[width=.5\textwidth]{images/4.png}}}
-    \caption{Линии уровня и точки испытаний ИАГП в двух синтетических задачах.}
+    \caption{Линии уровня и точки испытаний ИАГП в двух синтетических задачах}
     \label{fig:isolines}
@@
 
-\bibliographystyle{spmpsci}
-\bibliography{bibliography}{}
-%\begin{biblio}
-%\bibitem{eremin}
-%Еремин~И.И. Фейеровские методы для задач выпуклой и линейной оптимизации.
-%Челябинск: Изд-во ЮУрГУ, 2009. 200~с.
-%\bibitem{levin}
-%Левин~В.К. Отечественные суперкомпьютеры семейства МВС. URL:~\href{http://parallel.ru/mvs/levin.htm}{http://parallel.ru/mvs/levin.htm} (дата обращения: 27.05.2012).
-%\end{biblio}
+%\bibliographystyle{spmpsci}
+%\bibliography{bibliography}{}
+\begin{biblio}
+  \providecommand{\url}[1]{{#1}}
+  \providecommand{\urlprefix}{URL }
+  \expandafter\ifx\csname urlstyle\endcsname\relax
+    \providecommand{\doi}[1]{doi: \discretionary{}{}{}#1}\else
+    \providecommand{\doi}{doi: \discretionary{}{}{}\begingroup
+    \urlstyle{rm}\Url}\fi
+
+  \bibitem{BarkalovLebedev2017}
+  Barkalov K., Lebedev I.
+  \newblock Comparing two approaches for solving constrained global optimization
+    problems
+  \newblock // Learning and Intelligent Optimization. Springer International
+    Publishing, Cham. 2017. P. 301--306.
+  \newblock \doi{10.1007/978-3-319-69404-7_22}
+
+  \bibitem{BarkalovLebedev2017_2}
+  Barkalov K., Lebedev I.
+  \newblock Parallel algorithm for solving constrained global optimization
+    problems
+  \newblock // Parallel Computing Technologies. Springer International
+    Publishing, Cham. 2017. P. 396--404.
+  \newblock \doi{10.1007/978-3-319-62932-2_38}
+
+  \bibitem{BarkalovStrongin2018}
+  Barkalov K., Strongin R.
+  \newblock Solving a set of global optimization problems by the parallel
+    technique with uniform convergence //
+  \newblock J. of Global Optimization. 2018. Vol. 71, No. 1. P. 21--36.
+  \newblock \doi{10.1007/s10898-017-0555-4}
+
+  \bibitem{Beiranvand2017}
+  Beiranvand V., Hare W., Lucet Y.
+  \newblock Best practices for comparing optimization algorithms //
+  \newblock Optimization and Engineering. 2017. Vol. 18, No. 4. P. 815--848.
+  \newblock \doi{10.1007/s11081-017-9366-1}
+
+  \bibitem{Ehrgott2005}
+  Ehrgott M.
+  \newblock Multicriteria Optimization.
+  \newblock Springer-Verlag, Berlin, Heidelberg. 2005.
+  \newblock \doi{10.1007/3-540-27659-9}
+
+  \bibitem{Evtushenko2013}
+  Evtushenko Y., Posypkin M.
+  \newblock A deterministic approach to global box-constrained optimization //
+  \newblock Optim. Lett. 2013. Vol. 7 P. 819--829.
+  \newblock \doi{10.1007/s11590-012-0452-1}
+
+  \bibitem{Gaviano2003}
+  {Gaviano M., Kvasov D.E., Lera D., Sergeev Ya.D.}
+  \newblock Software for generation of classes of test functions with known local
+    and global minima for global optimization //
+  \newblock ACM Transactions on Mathematical Software. 2003. Vol. 29, No. 4. P.
+    469--480.
+  \newblock \doi{10.1145/962437.962444}
+
+  \bibitem{GergelBarkalov2019}
+  Gergel V., Barkalov K., Lebedev I., Rachinskaya M., Sysoyev A.
+  \newblock A flexible generator of constrained global optimization test
+    problems. 2019.
+  \newblock \doi{10.1063/1.5089976}
+
+  \bibitem{Jones2009}
+  {Jones D.R.}
+  \newblock The direct global optimization algorithm
+  \newblock // The Encyclopedia of Optimization. Springer, Heidelberg. 2009. P.
+    725--735.
+  \newblock \doi{10.1007/978-0-387-74759-0_128}
+
+  \bibitem{Paulavicius2011}
+  {Paulavivcius R., Zilinskas J., Grothey A.}
+  \newblock Parallel branch and bound for global optimization with combination of
+    lipschitz bounds //
+  \newblock Optim. Method. Softw.. 1997. Vol. 26, No. 3. P. 487--498.
+  \newblock \doi{10.1080/10556788.2010.551537}
+
+  \bibitem{RiquelmeLucken2015}
+  {Riquelme} N., {Von Lucken} C., {Baran} B.
+  \newblock Performance metrics in multi-objective optimization
+  \newblock // 2015 Latin American Computing Conference (CLEI). 2015. P. 1--11.
+  \newblock \doi{10.1109/clei.2015.7360024}
+
+  \bibitem{SergeyevKvasov2017}
+  Sergeyev Y., Kvasov D. Deterministic Global Optimization. 2017.
+  \newblock \doi{10.1007/978-1-4939-7199-2}
+
+  \bibitem{Sergeyev2013}
+  {Sergeyev Y.D., Strongin R.G., Lera D.} Introduction to global optimization
+    exploiting space-filling curves.
+  \newblock Springer Briefs in Optimization, Springer, New York. 2013.
+  \newblock \doi{10.1007/978-1-4614-8042-6}
+
+  \bibitem{Strongin2000}
+  {Strongin R.G., Sergeyev Ya.D.} Global optimization with non-convex
+    constraints. Sequential and parallel algorithms.
+  \newblock Kluwer Academic Publishers, Dordrecht. 2000.
+  \newblock \doi{10.1007/978-1-4615-4677-1}
+
+  \bibitem{BinhKorn1999}
+  To T.B., Korn B.
+  \newblock MOBES: A multiobjective evolution strategy for constrained
+    optimization problems. 1999.
+
+  \bibitem{grishaginClass}
+  {Гришагин В.А.}
+  \newblock {Операционные характеристики
+    некоторых алгоритмов глобального поиска}
+  \newblock // Проблемы статистической
+    оптимизации. Зинатне, Рига. 1978.
+
+  \bibitem{Norkin1992}
+  {Норкин В. И.}
+  \newblock О методе Пиявского для решения общей
+    задачи глобальной оптимизации //
+  \newblock Ж. вычисл. матем. и матем. физ. 1992. № 32 С.
+    992--1006.
+
+\end{biblio}
 

### 2020-02-06 2020/PAVT_Sovrasov/vm_journal/vm_paper.tex
PAVT: add a preliminary version for special issue

--- 2020-02-06 2020/PAVT_Sovrasov/pavt_paper.tex
@@
-\documentclass[11pt, oneside, a4paper]{article}
-%\usepackage[cp1251]{inputenc} % кодировка
-\usepackage[utf8]{inputenc} % кодировка
-\usepackage[english, russian]{babel} % Русские и английские переносы
-\usepackage{graphicx}          % для включения графических изображений
-\usepackage{cite}              % для корректного оформления литературы
-\usepackage{enumitem}
-\usepackage{pavt-ru}
+\documentclass{cmi}
+\usepackage{upgreek}
 
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{cite}
+% custom packages and commands
 \usepackage{doi}
 \usepackage{subfig}
-\usepackage[labelfont=bf,labelsep=period]{caption}
 
@@
 \DeclareMathOperator{\sign}{sign}
-\newtheorem{theorem}{Теорема}
 \renewcommand{\doitext}{DOI: }
@@
 
-% \title - название статьи
-% \authors - список авторов
+\classify{УДК 123.456, 789.012}
 
-\title{Параллельный алгоритм для получения равномерного приближения решений множества задач глобальной оптимизации с нелинейными ограничениями
-\footnote{Исследование выполнено при поддержке РНФ, проект №\,16-11-10150.}}
+\renewcommand*{\thefootnote}{\fnsymbol{footnote}} %to make a number in end of title delete this line
 
-\authors{В.В.~Соврасов, К.A.~Баркалов}
-\organizations{Нижегородский государственный университет им. Н.И.~Лобачевского}
+\title{ПАРАЛЛЕЛЬНЫЙ АЛГОРИТМ ДЛЯ ПОЛУЧЕНИЯ РАВНОМЕРНОГО ПРИБЛИЖЕНИЯ
+РЕШЕНИЙ МНОЖЕСТВА ЗАДАЧ ГЛОБАЛЬНОЙ ОПТИМИЗАЦИИ С НЕЛИНЕЙНЫМИ ОГРАНИЧЕНИЯМИ
+\footnote{Статья рекомендована к публикации программным комитетом Международной
+научной конференции <<Параллельные вычислительные технологии (ПаВТ) 2020>>.}%
+}
 
-% Аннотация заключается в окружение abstract
-\begin{abstract}
+\author{\copyright~2020~г. В.В.~Соврасов, К.A.~Баркалов}
+
+%\author{User1 \\ \href{mailto:me@somewhere.com}{me@somewhere.com}
+%	\and User2 \\ email %\href{mailto:someone@somewhere.com}{someone@somewhere.com} }
+
+\address{Нижегородский государственный университет им. Н.И.~Лобачевского\\
+  (603022 Нижний Новгород, пр.~Гагарина, д.~23, к.~2)}
+
+\email{E-mail: \href{mailto:sovrasov.vladislav@itmm.unn.ru}{sovrasov.vladislav@itmm.unn.ru},
+\href{mailto:konstantin.barkalov@itmm.unn.ru}{konstantin.barkalov@itmm.unn.ru}}
+
+\received{Поступила в редакцию: 26.02.2020}
+
+\maketitle{}
+
+\renewcommand*{\thefootnote}{\arabic{footnote}}
+\setcounter{footnote}{0}
+
+% оформление аннотации
+\begin{abstract}%
 В данной работе рассматривается построение параллельной версии алгоритма глобальной оптимизации, решающего одновременно множество задач с нелинейными ограничениями и получающего при этом равномерные оценки решений на этом множестве.
@@
 Эффективность реализованного алгоритма протестирована на наборах искусственно сгенерированных задач глобальной оптимизации.
+
+\keywords{глобальная оптимизация, параллельные вычисления, алгоритмы прямой оптимизации, равномерная сходимость}
 \end{abstract}
+\flushleft{\textbf{ОБРАЗЕЦ ЦИТИРОВАНИЯ}}
+\justifying
+\begin{citationplace}
+Соврасов~В.В., Баркалов~К.А. Параллельный алгоритм для получения равномерного приближения
+решений множества задач глобальной оптимизации с нелинейными ограничениями //
+Вестник ЮУрГУ. Серия: Вычислительная математика и информатика. 2020. Т.~X, №~Y. С.~Z1--Z2.
+DOI:~10.14529/cmseXXXXXX.
+\end{citationplace}
 
-\keywords{Глобальная оптимизация, параллельные вычисления, алгоритмы прямой оптимизации, равномерная сходимость}
-
-% \section{название} - заголовок раздела первого уровня
-% \subsection{название} - заголовок раздела второго уровня
-% \subsubsection{название} - заголовок раздела третьего уровня
-% Не используйте уровень вложенности заголовков больше трех!
-% Каждый абзац текста в статье начинается командой \par или пустой
-% строкой.
-
-\section{Введение}
-
+%Верхние колонтитулы
+\markboth{В.В.~Соврасов, К.A.~Баркалов}{Параллельный алгоритм для получения равномерного приближения\ldots}
+% Заголовки разделов формируются при помощи команд  \section{}, \subsection{}, \subsubsection{}
+\section*{Введение}
+\label{sec-intro}
 Нелинейная глобальная оптимизация невыпуклых функций традиционно считается одной из самых трудных
@@
     \subfloat[Решение внутри допустимой области]{
-    {\includegraphics[width=.5\textwidth]{images/2.png}}}
+    {\includegraphics[width=.5\textwidth]{pic/2.png}}}
     \subfloat[Решение на границе допустимой области]{
-    {\includegraphics[width=.5\textwidth]{images/4.png}}}
+    {\includegraphics[width=.5\textwidth]{pic/4.png}}}
     \caption{Линии уровня и точки испытаний ИАГП в двух синтетических задачах}
@@
     \subfloat[\(D_{max}\)]{
-    \label{fig:max_dev} {\includegraphics[width=.5\textwidth]{images/mixed_2d_max.pdf}}}
+    \label{fig:max_dev} {\includegraphics[width=.5\textwidth]{pic/mixed_2d_max.pdf}}}
     \subfloat[\(D_{avg}\)]{
-    \label{fig:avg_dev} {\includegraphics[width=.5\textwidth]{images/mixed_2d_avg.pdf}}}
+    \label{fig:avg_dev} {\includegraphics[width=.5\textwidth]{pic/mixed_2d_avg.pdf}}}
     \caption{Динамика величин \(D_{avg}\) и \(D_{max}\) в процессе решения множества двухмерных задач,
@@
     \subfloat[ИАГП, раздельное решение задач]{
-    \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{images/single_mco.pdf}}}
+    \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{pic/single_mco.pdf}}}
     \subfloat[ИАГП для множества задач]{
-    \label{fig:mco_pareto_2} {\includegraphics[width=.5\textwidth]{images/multi_mco.pdf}}}
+    \label{fig:mco_pareto_2} {\includegraphics[width=.5\textwidth]{pic/multi_mco.pdf}}}
     \caption{Численные оценки множества Парето в задаче (\ref{eq:mco_probem}), полученные после 2500 испытаний}
@@
 
-\section{Заключение}
+\section*{Заключение}
 
@@
 
-%\bibliographystyle{spmpsci}
-%\bibliography{bibliography}{}
+\vspace{1em}
+{\it Исследование выполнено при поддержке РНФ, проект №\,16-11-10150.}
+
 \begin{biblio}
-  \providecommand{\url}[1]{{#1}}
-  \providecommand{\urlprefix}{URL }
-  \expandafter\ifx\csname urlstyle\endcsname\relax
-    \providecommand{\doi}[1]{doi: \discretionary{}{}{}#1}\else
-    \providecommand{\doi}{doi: \discretionary{}{}{}\begingroup
-    \urlstyle{rm}\Url}\fi
+  \bibitem{BarkalovLebedev2017}
+  Barkalov K., Lebedev I.
+  \newblock Comparing two approaches for solving constrained global optimization
+    problems
+  \newblock // Learning and Intelligent Optimization. Springer International
+    Publishing, Cham. 2017. P. 301--306.
+  \newblock \doi{10.1007/978-3-319-69404-7_22}
+
+  \bibitem{BarkalovLebedev2017_2}
+  Barkalov K., Lebedev I.
+  \newblock Parallel algorithm for solving constrained global optimization
+    problems
+  \newblock // Parallel Computing Technologies. Springer International
+    Publishing, Cham. 2017. P. 396--404.
+  \newblock \doi{10.1007/978-3-319-62932-2_38}
+
+  \bibitem{BarkalovStrongin2018}
+  Barkalov K., Strongin R.
+  \newblock Solving a set of global optimization problems by the parallel
+    technique with uniform convergence //
+  \newblock J. of Global Optimization. 2018. Vol. 71, No. 1. P. 21--36.
+  \newblock \doi{10.1007/s10898-017-0555-4}
+
+  \bibitem{Beiranvand2017}
+  Beiranvand V., Hare W., Lucet Y.
+  \newblock Best practices for comparing optimization algorithms //
+  \newblock Optimization and Engineering. 2017. Vol. 18, No. 4. P. 815--848.
+  \newblock \doi{10.1007/s11081-017-9366-1}
+
+  \bibitem{Ehrgott2005}
+  Ehrgott M.
+  \newblock Multicriteria Optimization.
+  \newblock Springer-Verlag, Berlin, Heidelberg. 2005.
+  \newblock \doi{10.1007/3-540-27659-9}
+
+  \bibitem{Evtushenko2013}
+  Evtushenko Y., Posypkin M.
+  \newblock A deterministic approach to global box-constrained optimization //
+  \newblock Optim. Lett. 2013. Vol. 7 P. 819--829.
+  \newblock \doi{10.1007/s11590-012-0452-1}
+
+  \bibitem{Gaviano2003}
+  {Gaviano M., Kvasov D.E., Lera D., Sergeev Ya.D.}
+  \newblock Software for generation of classes of test functions with known local
+    and global minima for global optimization //
+  \newblock ACM Transactions on Mathematical Software. 2003. Vol. 29, No. 4. P.
+    469--480.
+  \newblock \doi{10.1145/962437.962444}
+
+  \bibitem{GergelBarkalov2019}
+  Gergel V., Barkalov K., Lebedev I., Rachinskaya M., Sysoyev A.
+  \newblock A flexible generator of constrained global optimization test
+    problems. 2019.
+  \newblock \doi{10.1063/1.5089976}
+
+  \bibitem{Jones2009}
+  {Jones D.R.}
+  \newblock The direct global optimization algorithm
+  \newblock // The Encyclopedia of Optimization. Springer, Heidelberg. 2009. P.
+    725--735.
+  \newblock \doi{10.1007/978-0-387-74759-0_128}
+
+  \bibitem{Paulavicius2011}
+  {Paulavivcius R., Zilinskas J., Grothey A.}
+  \newblock Parallel branch and bound for global optimization with combination of
+    lipschitz bounds //
+  \newblock Optim. Method. Softw.. 1997. Vol. 26, No. 3. P. 487--498.
+  \newblock \doi{10.1080/10556788.2010.551537}
+
+  \bibitem{RiquelmeLucken2015}
+  {Riquelme} N., {Von Lucken} C., {Baran} B.
+  \newblock Performance metrics in multi-objective optimization
+  \newblock // 2015 Latin American Computing Conference (CLEI). 2015. P. 1--11.
+  \newblock \doi{10.1109/clei.2015.7360024}
+
+  \bibitem{SergeyevKvasov2017}
+  Sergeyev Y., Kvasov D. Deterministic Global Optimization. 2017.
+  \newblock \doi{10.1007/978-1-4939-7199-2}
+
+  \bibitem{Sergeyev2013}
+  {Sergeyev Y.D., Strongin R.G., Lera D.} Introduction to global optimization
+    exploiting space-filling curves.
+  \newblock Springer Briefs in Optimization, Springer, New York. 2013.
+  \newblock \doi{10.1007/978-1-4614-8042-6}
+
+  \bibitem{Strongin2000}
+  {Strongin R.G., Sergeyev Ya.D.} Global optimization with non-convex
+    constraints. Sequential and parallel algorithms.
+  \newblock Kluwer Academic Publishers, Dordrecht. 2000.
+  \newblock \doi{10.1007/978-1-4615-4677-1}
+
+  \bibitem{BinhKorn1999}
+  To T.B., Korn B.
+  \newblock MOBES: A multiobjective evolution strategy for constrained
+    optimization problems. 1999.
+
+  \bibitem{grishaginClass}
+  {Гришагин В.А.}
+  \newblock {Операционные характеристики
+    некоторых алгоритмов глобального поиска}
+  \newblock // Проблемы статистической
+    оптимизации. Зинатне, Рига. 1978.
+
+  \bibitem{Norkin1992}
+  {Норкин В. И.}
+  \newblock О методе Пиявского для решения общей
+    задачи глобальной оптимизации //
+  \newblock Ж. вычисл. матем. и матем. физ. 1992. № 32 С. 992--1006.
+\end{biblio}
+
+% Сведения об авторах (Ф.И.О.(полностью), степень, звание, кафедра (отдел), вуз, e-mail)
+Владислав Валерьевич Соврасов, аспирант, Кафедра математического обеспечения и
+суперкомпьютерных технологий, Нижегородский государственный университет им. Н.И.~Лобачевского (Нижний Новгород, Российская Федерация)
+
+Константин александрович Баркалов, к.ф.м.н., доцент, Кафедра математического обеспечения и
+суперкомпьютерных технологий, Нижегородский государственный университет им. Н.И.~Лобачевского (Нижний Новгород, Российская Федерация)
+
+%\newpage
+
+{\phantom{eee} \hrule \vskip 4 pt}
+\classify{} % Код MSC не нужен
+%\classify{MSC 00000} % Коды MSC см. http://www.ams.org/msc/
+
+\title{\uppercase{Parallel global optimization algorithm for obtaining uniform convergence
+when simultaneously solving a set of global optimization problems}}
+
+\author{\copyright~2020 V.V.~Sovrasov, K.A.~Barkalov}
+
+\address{$^{1}$South Ural University (pr. Lenina 76, Chelyabinsk, 454080 Russia),\\$^{2}$Lomonosov Moscow State Universisty (GSP-1, Leninskie Gory 1, Moscow, 119991 Russia)}
+
+\email{E-mail: \href{mailto:sovrasov.vladislav@itmm.unn.ru}{sovrasov.vladislav@itmm.unn.ru},
+\href{mailto:konstantin.barkalov@itmm.unn.ru}{konstantin.barkalov@itmm.unn.ru}}
+
+\received{Received: 26.02.2020}
+
+\maketitle{}
+
+\begin{abstract}%
+  In this work building of a parallel version of a method simultaneously solving a set of
+  constrained global optimization problems is considered. The method is built on top of existing
+  one by extending it to problems with arbitrary nonlinear constraints and objective
+  satisfying the Lipschitz condition.
+
+  \keywordsen{global optimization, parallel computations.}
+\end{abstract}
+
+\begin{flushleft}
+  \textbf{FOR CITATION}
+\end{flushleft}
+\justifying
+
+\begin{citationplace}
+Sovrasov~V.V., Barkalov~K.A. The \LaTeXe{} Parallel global optimization algorithm for obtaining uniform convergence
+when simultaneously solving a set of global optimization problems.
+Bulletin of the South Ural State University.
+Series: Computational Mathematics and Software Engineering.
+2020.
+Vol.~X, no.~Y.
+P.~Z1--Z2.
+(in Russian) DOI: 10.14529/cmseXXXXXX.
+\end{citationplace}
+
+\begin{biblio_lat}
+%\itemsep -2pt
 
@@
   задачи глобальной оптимизации //
-  \newblock Ж. вычисл. матем. и матем. физ. 1992. № 32 С.
-    992--1006.
+\newblock Ж. вычисл. матем. и матем. физ. 1992. № 32 С. 992--1006.
 
-\end{biblio}
+\end{biblio_lat}
 

### 2020-02-07 2020/PAVT_Nested_Full/pavt_paper.tex
Before proof-reading

@@
 \begin{equation}\label{outer_prob}
-\varphi(\omega^*) = \min_\omega \varphi(\omega)
+\varphi(\omega^*) = \min_\omega \varphi(\omega),
 \end{equation}
@@
 
-%English begin
 Suppose $n\geq 1$  iterations of the method have been carried out, in which trials were performed at 
@@
 of $r$ allows to consider the product $r \mu$ as an estimate
-of the Lipschitz constant $L$ of the objective function .
+of the Lipschitz constant $L$ of the objective function.
 
@@
 convergence are given in \cite{Barkalov2018,Sergeyev2013,Strongin2018,Strongin2000}.
-%English end
+%Замечание по поводу теоретических свойств и сравнения алгоритмов
+The results of the experimental comparison  (see, for example, \cite{Sovrasov2019}) show that global search algorithm is superior to many known methods of similar purpose, including both deterministic and heuristic  algorithms.
 
@@
 
+% В последней серии экспериментов было решено 10 задач, в которых в качестве многоэкстремальной части использовались трехмерные задачи из класса GKLS.
+Данный класс задач часто используется для тестирования методов глобальной оптимизации [ссылки].
+Размерность локальных подзадач была равна N = 100.
+Задачи решались на 1, 8, 16 и 32 узлах кластера, на которых использовалось от 4 до 128 MPI-процессов.
+Параметры запуска такие же, как в предыдущих сериях экспериментов.
+Результаты решения представлены в таблице, ускорение посчитано относительно запуска на одном узле на 4 mpi процесса.
+
+In the last series of experiments, 10 problems were solved, in which three-dimensional functions from GKLS class \cite{Gaviano2003} were used as a multiextremal part. 
+This class of functions is often used for investigating global optimization algorithms \cite{Sergeyev2015,Paulavicius2014,Barkalov2018}.
+The dimensionality of the local subproblems was equal $N=100$.
+The problems were solved at 1, 8, 16 and 32 cluster nodes, on which from 4 to 128 MPI processes were used.
+The running parameters were the same as in the previous series of experiments.
+The results of solving are presented in Table \ref{tab3}, where speedup is calculated relative to running on a single node using 4 MPI processes.
+
+%In the last series of experiments, 10 problems of the kind (\ref{test_problem}) were solved, in which the dimensionality of the local subproblems was equal $N=100$.
+The running parameters were the same as in the previous series of experiments.
+The results of solving are presented in Table \ref{tab3}.
+
+\begin{table}
+\centering
+\caption{Results of solving a series of problems of dimensionality $N=103$}\label{tab3}
+\begin{tabular}{cccc}
+\hline\noalign{\smallskip}
+ $\;\;\;p\;\;\;$  &  $\;\;\;K_{av}\;\;\;$ &  $\;\;\;T_{av}\;\;\;$ & $\;\;\;S\;\;\;$ \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+ 4  & 730.8  & 63.4 & ---  \\
+ 32 & 87.4   & 7.37 & 8.6 \\
+ 64 & 37.0   & 3.77 & 16.8 \\
+ 128 & 16.5   & 1.78 & 35.7 \\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
 \section{Conclusion}
@@
 In the case of solving  the problems with more complex local subproblems (with the number of the 
-local variables $N=100$), the speedup increased up to 27.4 when using 32 processes.
+local variables $N=100$), the speedup increased up to 35 when using 128 processes. In the last case the speedup was calculated with respect to the use of 4 processes on one node.
 The developed scheme of parallel computations can be applied in the identification of complex 

### 2020-02-07 2020/PAVT_Sovrasov/vm_journal/vm_paper.tex
PAVT: update English in the special issue version

@@
 
-\address{$^{1}$South Ural University (pr. Lenina 76, Chelyabinsk, 454080 Russia),\\$^{2}$Lomonosov Moscow State Universisty (GSP-1, Leninskie Gory 1, Moscow, 119991 Russia)}
+\address{Lobachevsky State University of Nizhni Novgorod, Russia (pr. Gagarina 23(2), Nizhni Novgorod, 603022 Russia)}
 
@@
 \maketitle{}
-
 \begin{abstract}%
   In this work building of a parallel version of a method simultaneously solving a set of
-  constrained global optimization problems is considered. The method is built on top of existing
-  one by extending it to problems with arbitrary nonlinear constraints and objective
-  satisfying the Lipschitz condition.
+  constrained global optimization problems is considered. This method converges uniformly
+  to solutions of all the problems. That allows the method to arrange computational resources in an
+  optimal way, since uniform convergence guarantees approximately equal precision
+  of numerical solutions at the whole set of problems at the each iteration of optimization.
+  Sets of similar global optimization problems appear for an instance after scalarization
+  of multi-objective problems or when a global optimization problem has a discrete parameter
+  which takes a finite number of possible values. The considered method uses Peano-type curves
+  to transform multidimensional problems into univariate ones. Efficiency of the implemented
+  parallel algorithm is evaluated on several sets of synthetically generated constrained global optimization problems.
 
-  \keywordsen{global optimization, parallel computations.}
+  \keywordsen{global optimization, parallel computations, derivative-free optimization, uniform convergence}
 \end{abstract}
@@
 \begin{citationplace}
-Sovrasov~V.V., Barkalov~K.A. The \LaTeXe{} Parallel global optimization algorithm for obtaining uniform convergence
+Sovrasov~V.V., Barkalov~K.A. Parallel global optimization algorithm for obtaining uniform convergence
 when simultaneously solving a set of global optimization problems.
@@
   lipschitz bounds //
-\newblock Optim. Method. Softw.. 1997. Vol. 26, No. 3. P. 487--498.
+\newblock Optim. Method. Softw. 1997. Vol. 26, No. 3. P. 487--498.
 \newblock \doi{10.1080/10556788.2010.551537}
@@
 \bibitem{grishaginClass}
-{Гришагин В.А.}
-\newblock {Операционные характеристики
-  некоторых алгоритмов глобального поиска}
-\newblock // Проблемы статистической
-  оптимизации. Зинатне, Рига. 1978.
+{Grishagin V.}
+\newblock {Operating characteristics of some global search algorithms}
+\newblock // Problems of Stochastic Search. Zinatne, Riga. 1978. (in Russian)
 
 \bibitem{Norkin1992}
-{Норкин В. И.}
-\newblock О методе Пиявского для решения общей
-  задачи глобальной оптимизации //
-\newblock Ж. вычисл. матем. и матем. физ. 1992. № 32 С. 992--1006.
+{Norkin V. I.}
+\newblock Towards Pijavskyj's method for solving common global optimization problem //
+\newblock J. Comput. Math. Math. Phys. 1992. Vol. 32, no. 7. P. 992--1006. (in Russian)
 

### 2020-02-10 2020/PAVT_Nested_Full/pavt_paper.tex
After proof-reading of introduction

@@
 
-In the present paper, the global optimization problems and the numerical methods of solving these 
-ones are considered. 
-Regarding the character of the dependence of the objective function on its parameters, an assumption 
-is made that it is multiextremal with respect to selected variables only, and the dependence on the rest 
-ones has a local character. 
-The problems of this type may arise in the identification of the unknown parameters of the 
-mathematical models on the base of the results of experiments. 
-A parallel computational scheme, which accounts for this feature has been proposed.
-The novel scheme is based on the idea of the nested (recursive) optimization when at the upper 
-recursion level the optimization with respect to the globally-meaning variables is performed (using 
-the global optimization method), and at the lower one the local optimization problems are solved (using 
-the local method). The generated local subproblems will not affect each other, and solving of these 
-ones can be performed in parallel. The numerical experiments have been conducted on several 
-hundred test problems confirming the efficiency of the proposed scheme of parallel computations.
+The present paper considers global optimization problems and the numerical methods of solving them.
+An assumption is made regarding the dependence of the objective function on its parameters: it is multiextremal with respect to selected variables only, and the dependence on the other variables has a local character.
+Problems of this type may arise in identifying the unknown parameters of the mathematical models, based on experimental results.
+A parallel computational scheme, which accounts for this feature, is being proposed.
+This novel scheme is based on the idea of nested (recursive) optimization, when at the upper recursion level the optimization with respect to the significant variables is performed (using the global optimization method), and at the lower one local optimization problems are solved (using the local method).
+The generated local subproblems will not affect each other, and solving these problems can be performed in parallel.
+Numerical experiments have been conducted on several hundred test problems, confirming the efficiency of the proposed scheme of parallel computations.
 
@@
 
-The global optimization problems arise often in various fields of modern science and technology. For 
-example, the problems of finding the configurations of various chemical compounds corresponding to 
-the minimum interaction energy are relevant \cite{Posypkin2014}. In particular,  such problems arise 
+Global optimization problems arise frequently in various fields of modern science and technology.
+For example, problems of finding the configurations of various chemical compounds corresponding to the minimum interaction energy are relevant \cite{Posypkin2014}.
+In particular,  such problems arise
 in the development of new medical drugs \cite{Sulimov}. 
 Another example is the problem of finding optimal control parameters for complex processes \cite{Kalyulin2017,Modorskii2016}.
-The global optimization problems arise in many other applications as well (see the review \cite{Pinter2006}).
+Global optimization problems arise in many other applications as well (see the review \cite{Pinter2006}).
 
-The application of the global optimization methods for the identification of the parameters of the 
-mathematical models based on the data of experiments became a classical one. In the problems of this 
-type, it is required to carry out a search for the values of the unknown parameters of the model, at 
-which the results of computations according to the model are close to the results obtained 
-experimentally.
+Application of global optimization methods for identifying the parameters of the mathematical models based on the data of experiments has become a classical practice.
+Problems of this type require a search for the values of the unknown parameters of the model, from which computational results according to the model are close to the results obtained experimentally.
 
-The number of parameters, which are required to be identified in such a way may be tens and hundreds 
-in complex mathematical models \cite{Akhmadullina2017,Nurislamova2016}. In the problems of 
-such dimensionality, the regular methods of search for the global solutions cannot be applied because 
-of extremely large computational costs for covering the search domain by the trial points. This 
-remains to be true even in the case of the use of efficient algorithms (for example, 
-\cite{Evtushenko2009,Jones2009,Paulavicius2011}) constructing essentially nonuniform coverages.
+The number of parameters which are required to be identified in such a way, may reach tens and hundreds in complex mathematical models \cite{Akhmadullina2017,Nurislamova2016}.
+In problems of such dimensionality, regular search methods for global solution cannot be applied because of the extremely large computational costs for covering the search domain by the trial points.
+This remains true even with the use of efficient algorithms (for example, \cite{Evtushenko2009,Jones2009,Paulavicius2011}) constructing essentially nonuniform coverages.
 
-However, in the problems of the model identification, the parameters, which will affect the result 
-mostly can be selected often. The number of such parameters is small as a rule. The rest parameters 
-either cause an insufficient effect (in this case, the multiextremality can be neglected) or the 
-dependence on the second group of parameters have a local character.
+However, in the problems of model identification, the parameters, which will affect the result most significantly, can often be selected.
+As a rule, the number of such parameters is small.
+The rest of the parameters either cause an insufficient effect (in this case, multiextremality can be neglected) or dependence on the second group of parameters has a local character.
 
-So far, the development of the parallel algorithms for the minimization of the essentially 
-multidimensional functions (hundreds of variables), which account for different character of 
-dependence of the objective function on different groups of parameters is relevant. Namely, the 
-objective function is assumed to be multiextremal with respect to a part of variables only. The rest 
-variables have a local effect, i.e. the objective function is unimodal with respect to these ones. 
-In this case, the solving of the problem can be organized according to the nested (recursive) 
-optimization scheme. The solving of the multiextremal subproblem, for which the use of complex 
-global search algorithms is required, will be performed at the upper level of recursion.
-The unimodal subproblems (each corresponding to a fixed set of values of the first part of 
-parameters) will be solved at the lower level. For solving these subproblems, the efficient 
-methods of local optimization, linear programming, or linear algebra can be applied (subject to the 
-character of the local subproblems).
+So far, the development of parallel algorithms for minimizing the essentially multidimensional functions (hundreds of variables), which accounts for the dependence of the objective function on different groups of parameters, is relevant.
+Specifically, the objective function is assumed to be multiextremal with respect to some of the variables.
+The rest of the variables have a local effect, i.e., the objective function is unimodal with respect to them.
+In this case, solving the problem can be organized according to the nested (recursive) optimization scheme.
+Solving the multiextremal subproblem, which requires the use of complex global search algorithms, will be performed at the upper level of recursion.
+The unimodal subproblems (each corresponding to a fixed set of values of the first part of parameters) will be solved at the lower level.
+For solving these subproblems, efficient methods of local optimization, linear programming, or linear algebra can be applied (subject to the character of the local subproblems).
 
-The paper text describing the results of the present study is organized as follows. 
-In Section 2, the description of the nested optimization scheme taking into account different properties 
-of the objective function at different nesting level is given. Particular example of a problem of this 
-class is presented. A general scheme of organization of the parallel computations is described. 
-Section 3 is devoted to the description of the parallel global search algorithm using the space-filling 
-curves. The computational scheme of the parallel algorithm is presented, its main properties are 
-described.
-In Section 4, the methods of solving the local subproblems within the framework of the nested 
-optimization scheme are discussed. A brief description of the properties of the local extremum search 
-problems and known methods of solving these ones is given. 
-Section 5 contains the results of numerical experiment. Here, the evaluation of the scalability of the 
-proposed parallel computation scheme in solving a series of test problems is performed.  
+The text describing the results of the present study is organized as follows.
+Section 2 describes the nested optimization scheme taking into account different properties of the objective function at different nesting levels.
+A particular example of a problem of this class is presented, along with a general scheme of the organization of parallel computations.
+Section 3 is devoted to the description of the parallel global search algorithm using space-filling curves.
+The computational scheme of the parallel algorithm is presented, and its main properties are described.
+Section 4 discusses the methods of solving the local subproblems within the framework of the nested optimization scheme.
+A brief description of the properties of the local extremum search problems and known methods of solving them is given.
+Section 5 contains the results of a numerical experiment.
+The evaluation of the scalability of the proposed parallel computation scheme in solving a series of test problems was performed.
 Section 6 concludes the paper.
@@
 In the last series of experiments, 10 problems were solved, in which three-dimensional functions from GKLS class \cite{Gaviano2003} were used as a multiextremal part. 
-This class of functions is often used for investigating global optimization algorithms \cite{Sergeyev2015,Paulavicius2014,Barkalov2018}.
+This class of functions is often used for investigating global optimization algorithms \cite{Barkalov2018,Paulavicius2014,Sergeyev2015}.
 The dimensionality of the local subproblems was equal $N=100$.
@@
 The running parameters were the same as in the previous series of experiments.
-The results of solving are presented in Table \ref{tab3}, where speedup is calculated relative to running on a single node using 4 MPI processes.
+The results of solving are presented in Table \ref{tab3}, where speedup is calculated with respect to  running on a single node using 4 MPI processes.
 
@@
 
-The results of experiments performed on the series of the optimization problems
-%, in which a multiextremal part and a unimodal one can be selected,
-demonstrated the scheme of the 
-parallel computations proposed in the present work to have a good potential for parallelization. 
-In particular, when solving a series of 100 problems with total number of variables equal to 52 (from 
-which, two first variables cause a global effect on the objective function and fifty affected this one 
-locally), a speedup of 24.6 has been obtained when using  32 processes. 
-In the case of solving  the problems with more complex local subproblems (with the number of the 
-local variables $N=100$), the speedup increased up to 35 when using 128 processes. In the last case the speedup was calculated with respect to the use of 4 processes on one node.
-The developed scheme of parallel computations can be applied in the identification of complex 
-mathematical models, which are featured by a large number (tens and hundreds) of unknown 
-parameters.
-
+Results of experiments performed on the series of optimization problems demonstrate that the scheme of parallel computations proposed in the present work has good potential for parallelization.
+Specifically, when solving a series of 100 problems with a total number of variables equal to 52 (from which the two first variables caused a global effect on the objective function and fifty affected the one locally), a speedup of 24.6 was obtained when using 32 processes.
+When solving  problems with more complex local subproblems (with the number of the local variables N=100), the speedup increased up to 35 when using 128 processes.
+In the last case the speedup was calculated with respect to the use of 4 processes on one node.
+The developed scheme of parallel computations can be applied in identifying complex mathematical models, which are featured by large numbers (tens and hundreds) of unknown parameters.
 %

### 2020-02-13 2020/PAVT_Nested_Full/pavt_paper.tex
emails of all authors

@@
 %
-\title{Combining of local and global search in the parallel nested optimization scheme 
-\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
+\title{Combining of local and global search in the parallel nested optimization scheme\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
 }
@@
 %%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{Konstantin Barkalov and Ilya Lebedev, and Maria Kocheganova}
+\tocauthor{Konstantin Barkalov and Ilya Lebedev and Maria Kocheganova and Victor Gergel}
 %
 \institute{Lobachevsky State University of Nizhni Novgorod, Russia  \\
-	\email{konstantin.barkalov@itmm.unn.ru}
+	\email{konstantin.barkalov@itmm.unn.ru},
+	\email{ilya.lebedev@itmm.unn.ru},
+	\email{maria.rachinskaya@itmm.unn.ru},
+	\email{gergel@unn.ru}
 }

### 2020-02-26 2019/Automation/Barkalov.tex
1.5 interval
text corrections

@@
 
+\usepackage{setspace}
+\onehalfspacing % полуторный интервал для всего текста
+
 \begin{document}  %%%!!!
 
-\year{2019}
-\title{АДАПТИВНАЯ ГЛОБАЛЬНАЯ ОПТИМИЗАЦИЯ НА ОСНОВЕ БЛОЧНО-РЕКУРСИВНОЙ СХЕМЫ РЕДУКЦИИ РАЗМЕРНОСТИ}%
+\year{2020}
+\title{АДАПТИВНАЯ ГЛОБАЛЬНАЯ ОПТИМИЗАЦИЯ НА ОСНОВЕ БЛОЧНО-РЕКУРСИВНОЙ СХЕМЫ РЕДУКЦИИ РАЗМЕРНОСТИ }%
 \thanks{Исследование выполнено за счет гранта Российского научного фонда (проект \mbox{№\,16-11-10150}).}
@@
 \authors{
-Р.Г.~СТРОНГИН, д-р~физ.-мат.~наук\\
-В.П.~ГЕРГЕЛЬ, д-р~техн.~наук\\
-К.А.~БАРКАЛОВ, канд.~физ.-мат.~наук\\
+Р.Г.~СТРОНГИН, д-р~физ.-мат.~наук (strongin@unn.ru)\\
+В.П.~ГЕРГЕЛЬ, д-р~техн.~наук (gergel@unn.ru)\\
+К.А.~БАРКАЛОВ, канд.~физ.-мат.~наук (barkalov@vmk.unn.ru)\\
 (Нижегородский государственный университет им. Н.И. Лобачевского)}
@@
 \begin{abstract}
-В статье рассматриваются задачи многомерной многоэкстремальной оптимизации и численные методы их решения.
+Рассматриваются задачи многомерной многоэкстремальной оптимизации и численные методы их решения.
 Об оптимизируемой функции делается лишь общее предположение, что она удовлетворяет условию Липшица с априори неизвестной константой (задачи такого типа часто встречаются в приложениях).
-Рассмотрено два способа редукции размерности в задачах многомерной оптимизации: использованные кривых Пеано (разверток) и рекурсивная многошаговая схема.
+Рассмотрено два способа редукции размерности в задачах многомерной оптимизации: использование кривых Пеано (разверток) и рекурсивная многошаговая схема.
 Предложена обобщенная схема, комбинирующая эти два подхода.
-В новой схеме решение многомерной задачи сводится к решению семейства задач меньшей размерности, в которых, в свою очередь, используются развертки.
+В новой схеме решение многомерной задачи сводится к решению семейства задач меньшей размерности, в которых в свою очередь используются развертки.
 Реализован адаптивный алгоритм, в котором все возникающие подзадачи решаются одновременно.
@@
 & \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
-& D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
+& D=\left\{y\in \text{R}^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
 \end{eqnarray}
-Предполагается, что целевая функция может быть многоэкстремальной, задана неявно (функция вида <<черный ящик>>), а вычисление ее значений связано с решением задачи численного моделирования и является трудоемкой операцией.
+Предполагается, что целевая функция может быть многоэкстремальной, задана неявно (функция вида ``черный ящик''), а вычисление ее значений связано с решением задачи численного моделирования и является трудоемкой операцией.
 
-Любая возможность достоверно оценить глобальный оптимум в многоэкстремальной задаче с функциями вида <<черный ящик>> принципиально основана на априорной информации, позволяющей связать возможные значения целевой функции с ее известными значениями в точках проведенных испытаний.
-Для многих прикладных задачах типичной является ситуация, когда ограниченное изменение вектора параметров $y$ вызывает ограниченное изменение значений $\varphi(y)$.
+Любая возможность достоверно оценить глобальный оптимум в многоэкстремальной задаче с функциями вида ``черный ящик'' принципиально основана на априорной информации, позволяющей связать возможные значения целевой функции с ее известными значениями в точках проведенных испытаний.
+Для многих прикладных задач типичной является ситуация, когда ограниченное изменение вектора параметров $y$ вызывает ограниченное изменение значений $\varphi(y)$.
 Математической моделью, описывающей указанное предположение, является условие Липшица
@@
 Предположение липшицевости целевой функции типично для многих подходов к разработке оптимизационных алгоритмов.
-Первые методы липшицевой оптимизации были предложены в начале 70-х годов \cite{Evtushenko71,Pijavski72,Shubert72}; с тех пор данное направление продолжает активно развиваться \cite{Jones93,Pinter96,Zilinskas08,Evtushenko07,Evtushenko09}.
+Первые методы липшицевой оптимизации были предложены в начале 70-х гг.
+$\overline{\underline{\text{XX}}}$ в.
+\cite{Evtushenko71,Pijavski72,Shubert72}; с тех пор данное направление продолжает активно развиваться \cite{Jones93,Pinter96,Zilinskas08,Evtushenko07,Evtushenko09}.
 Например, многие известные методы основаны на различных способах разбиения области поиска на систему подобластей и последующего выбора наиболее перспективной подобласти для размещения очередного испытания (вычисления целевой функции).

### 2020-02-26 2019/Automation/Barkalov.tex
correct up to page 7

@@
 \documentclass[12pt]{a&t}
-\usepackage{hyperref}
+%\usepackage{hyperref}
 \usepackage{graphicx}
+\usepackage{soul}
 
@@
 \]
-%Отметим, что значение константы $L$ обычно является априори неизвестным, что делает ее оценку одной из ключевых проблем при построении методов липшицевой оптимизации.
+%Отметим, что значение константы $L$ обычно является априори неизвестным, что делает ее оценку одной из ключевых проблем при построении методов липшицевой оптимизации.$\overline{\underline{\text{XX}}}$
 Предположение липшицевости целевой функции типично для многих подходов к разработке оптимизационных алгоритмов.
 Первые методы липшицевой оптимизации были предложены в начале 70-х гг.
-$\overline{\underline{\text{XX}}}$ в.
+XX в.
 \cite{Evtushenko71,Pijavski72,Shubert72}; с тех пор данное направление продолжает активно развиваться \cite{Jones93,Pinter96,Zilinskas08,Evtushenko07,Evtushenko09}.
 Например, многие известные методы основаны на различных способах разбиения области поиска на систему подобластей и последующего выбора наиболее перспективной подобласти для размещения очередного испытания (вычисления целевой функции).
-Результаты, полученные в данном направлении, представлены в работах \cite{Jones09,Zilinskas10,Evtushenko13,Kvasov13,Paulavicius16}.
+Результаты, полученные в данном направлении, представлены в публикациях \cite{Jones09,Zilinskas10,Evtushenko13,Kvasov13,Paulavicius16}.
 
-В настоящее время для решения задач оптимизации с функциями вида <<черный ящик>> широко используются генетические и популяционные алгоритмы (см., например, \cite{Kureichik04,Karpenko08}), которые так или иначе основаны на идеях случайного поиска.
+В настоящее время для решения задач оптимизации с функциями вида ``черный ящик'' широко используются генетические и популяционные алгоритмы (см., например, \cite{Kureichik04,Karpenko08}), которые так или иначе основаны на идеях случайного поиска.
 В силу простоты реализации и использования они получили большую популярность, однако по качеству работы (численной оценкой которого может служить число корректно решенных задач из некоторого набора) они существенно уступают детерминированным алгоритмам \cite{Kvasov18,Sergeyev18}.
@@
 
-Разработанные методы основаны на редукции исходной многомерной задачи к эквивалентной одномерной или системе одномерных подзадач и последующим решением одномерных задач эффективными методами оптимизации функций одной переменной.
+Разработанные методы основаны на редукции исходной многомерной задачи к эквивалентной одномерной или к системе одномерных подзадач с последующим решением одномерных задач эффективными методами оптимизации функций одной переменной.
 Предложено две такие схемы: редукция на основе кривых, заполняющих пространство (кривых Пеано, или \textit{разверток}) \cite{Strongin91,Strongin00}, и схема рекурсивной вложенной оптимизации (\textit{многошаговая схема}) \cite{Grishagin97,Grishagin01}.
-В статье \cite{Grishagin16} предложена адаптивная многошаговая схема, существенно повышающая эффективность оптимизации по сравнению с базовым прототипом.
-В данной работе предлагается обобщение адаптивной схемы редукции размерности, комбинирующее использование вложенной оптимизации и кривых Пеано.
+В \cite{Grishagin16} предложена адаптивная многошаговая схема, существенно повышающая эффективность оптимизации по сравнению с базовым прототипом.
+В данной статье предлагается обобщение адаптивной схемы редукции размерности, комбинирующее использование вложенной оптимизации и кривых Пеано.
 При таком подходе вложенные подзадачи в адаптивной схеме могут быть как одномерными, так и многомерными; в последнем случае для редукции размерности вложенных подзадач используются развертки.
@@
 
-В соответствии с алгоритмом, первые два испытания проводятся в граничных точках интервала $[0,1]$, т.е. $x^0=0,\;x^1=1$. 
-В этих точка вычисляются значения целевой функции $z^0=\varphi(x^0),\;z^1=\varphi(x^1)$, и устанавливается значение счетчика $k=1$. 
+В соответствии с алгоритмом первые два испытания проводятся в граничных точках отрезка $[0,1]$, т.е. $x^0=0,\;x^1=1$. 
+В этих точках вычисляются значения целевой функции $z^0=\varphi(x^0),\;z^1=\varphi(x^1)$ и устанавливается значение счетчика $k=1$. 
 Точка очередного испытания $x^{k+1}, k\geq 1,$ выбирается в соответствии со следующими правилами.
 
-\textbf{Шаг 1.} Перенумеровать нижним индексом (начиная с 0) точки $x^i,\:0\leq i\leq k$, проведенных испытаний в порядке возрастания координаты, т.е.,
+\so{\sl Шаг 1}. Перенумеровать нижним индексом (начиная с 0) точки $x^i,\:0\leq i\leq k$, проведенных испытаний в порядке возрастания координаты, т.е.
 \[
@@
 \] 
-Сопоставить точкам $x_i, 0\leq i\leq k$, вычисленные в них значения целевой функции $z_i=\varphi(x_i), 0\leq i\leq k$.
+Сопоставить точкам $x_i, \; 0\leq i\leq k$, вычисленные в них значения целевой функции $z_i=\varphi(x_i), \; 0\leq i\leq k$.
 
-\textbf{Шаг 2.} Вычислить максимальное абсолютное значение относительной первой разности
+\so{\sl Шаг 2}. Вычислить максимальное абсолютное значение относительной первой разности
 \begin{equation}\label{mu}
-\mu=\max_{1\leq i\leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i}
+\mu=\max_{1\leq i\leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
 \end{equation}
-где $\Delta_i = x_i-x_{i-1}$. Если вычисленное в соответствии с данной формулой значение равно 0, то положить $\mu = 1$.
+где $\Delta_i = x_i-x_{i-1}$. Если вычисленное в соответствии с данной формулой значение равно нулю, то положить $\mu = 1$.
 
-\textbf{Шаг 3.} Для всех интервалов $(x_{i-1},x_i),1\leq i\leq k$,  вычислить значение
+\so{\sl Шаг 3}. Для всех интервалов $(x_{i-1},x_i),1\leq i\leq k$,  вычислить значение
 \begin{equation}\label{R}
@@
 
-\textbf{Шаг 4.} Найти интервал $(x_{t-1},x_t)$ с максимальной характеристикой
+\so{\sl Шаг 4}. Найти интервал $(x_{t-1},x_t)$ с максимальной характеристикой
 \begin{equation}\label{MaxR}
@@
 
-\textbf{Шаг 5.} Провести новое испытание в точке
+\so{\sl Шаг 5}. Провести новое испытание в точке
 \begin{equation}\label{xk1}
@@
 \begin{equation}
-\left\{y(x):0\leq x \leq 1 \right\} = \left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq N\right\}.
+\left\{y(x):0\leq x \leq 1 \right\} = \left\{y\in \text{R}^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq N\right\}.
 \end{equation}
@@
 (см. \cite{Strongin00}). Если функция $\varphi(y)$ в области $D$ удовлетворяла условию Липшица, тогда редуцированная функция  
-$\varphi(y(x)), x\in [0,1]$ будет удовлетворять равномерному условию Гельдера 
+$\varphi(y(x)), x\in [0,1]$, будет удовлетворять равномерному условию Гельдера 
 \begin{equation}\label{Holder}
@@
 
-Условия (\ref{Holder}), (\ref{HL}) позволяет легко обобщить <<одномерный>> алгоритм из раздела \ref{SectionCore} для решения многомерных задач. 
-Для этого достаточно заменить длины интервалов $\Delta_i$ в формулах (\ref{mu}),(\ref{R}) на длины 
+Условия (\ref{Holder}), (\ref{HL}) позволяют легко обобщить ``одномерный'' алгоритм из раздела \ref{SectionCore} для решения многомерных задач. 
+Для этого достаточно заменить длины интервалов $\Delta_i$ в формулах (\ref{mu}), (\ref{R}) на длины 
 \begin{equation}
@@
 \end{equation}
-Вычисление значений функции $\varphi_2$, в свою очередь, требует одномерной минимизации функции $\varphi_3$, и т.д. вплоть до
+Вычисление значений функции $\varphi_2$ в свою очередь требует одномерной минимизации функции $\varphi_3$ и т.д. вплоть до
 решения задачи
 \begin{equation}
-\varphi_{N-1}(y_1,...,y_{N-1}) = \min_{ y_{N} \in\left[a_{N},b_{N}\right]} \varphi_{N}(y_1,...,y_{N}),
+\varphi_{N-1}(y_1,...,y_{N-1}) = \min_{ y_{N} \in\left[a_{N},b_{N}\right]} \varphi_{N}(y_1,...,y_{N})
 \end{equation}
@@
   \includegraphics[width=0.7\textwidth]{fig1.pdf} 
-  \caption{Взаимосвязи между подзадачами в многошаговой схеме}\label{fig0} 
+  \caption{Взаимосвязи между подзадачами в многошаговой схеме.}\label{fig0} 
 \end{center}
@@
 где $i$-я блочная переменная, т.е. вектор $u_i$, состоит из взятых последовательно компонент вектора $y$, т.е. $u_1=(y_1,y_2,...,y_{N_1})$, $u_2=(y_{N_1+1},y_{N_1+2}
-,...,y_{N_1+N_2})$,..., $u_M=(y_{N-N_M+1},y_{N-N_M+2},...,y_{N})$, а $N_1+
-N_2+...+N_M=N$.
+,...,y_{N_1+N_2})$, ..., $u_M=(y_{N-N_M+1}, y_{N-N_M+2},..., y_{N})$, а $N_1+N_2+...+N_M=N$.
 
@@
 \begin{equation}\label{block_nested_i}
-\varphi_i(u_1,...,u_i) = \min_{u_{i+1}\in D_{i+1}} \varphi_{i+1}(u_1,...,u_i,u
-_{i+1}), 1\leq i\leq M-1.
+\varphi_i(u_1,...,u_i) = \min_{u_{i+1}\in D_{i+1}} \varphi_{i+1}(u_1,...,u_i,u_{i+1}), \; 1\leq i\leq M-1,
 \end{equation}
@@
 Для случая одномерных подзадач данный подход был теоретически обоснован и апробирован в \cite{Grishagin16,Grishagin2016_1,Grishagin18}.
-Настоящая работа предлагает обобщение адаптивной схемы для случая многомерных подзадач.
+Настоящая статья предлагает обобщение адаптивной схемы для случая многомерных подзадач.
 

### 2020-02-27 2020/PAVT_Sovrasov/vm_journal/vm_paper.tex
PAVT: Final fixes to the journal version

@@
 
-\classify{УДК 123.456, 789.012}
+\classify{УДК 004.021, 519.6}
 
@@
 
-\received{Поступила в редакцию: 26.02.2020}
+\received{Поступила в редакцию: 27.02.2020}
 
@@
 
-\received{Received: 26.02.2020}
+\received{Received: 27.02.2020}
 
@@
 \newblock Comparing two approaches for solving constrained global optimization
-  problems
-\newblock // Learning and Intelligent Optimization. Springer International
+  problems.
+\newblock Learning and Intelligent Optimization. Springer International
   Publishing, Cham. 2017. P. 301--306.
@@
 \newblock Parallel algorithm for solving constrained global optimization
-  problems
-\newblock // Parallel Computing Technologies. Springer International
+  problems.
+\newblock Parallel Computing Technologies. Springer International
   Publishing, Cham. 2017. P. 396--404.
@@
 \newblock Solving a set of global optimization problems by the parallel
-  technique with uniform convergence //
+  technique with uniform convergence.
 \newblock J. of Global Optimization. 2018. Vol. 71, No. 1. P. 21--36.
@@
 Beiranvand V., Hare W., Lucet Y.
-\newblock Best practices for comparing optimization algorithms //
+\newblock Best practices for comparing optimization algorithms.
 \newblock Optimization and Engineering. 2017. Vol. 18, No. 4. P. 815--848.
@@
 Evtushenko Y., Posypkin M.
-\newblock A deterministic approach to global box-constrained optimization //
+\newblock A deterministic approach to global box-constrained optimization.
 \newblock Optim. Lett. 2013. Vol. 7 P. 819--829.
@@
 \newblock Software for generation of classes of test functions with known local
-  and global minima for global optimization //
+  and global minima for global optimization.
 \newblock ACM Transactions on Mathematical Software. 2003. Vol. 29, No. 4. P.
@@
 {Jones D.R.}
-\newblock The direct global optimization algorithm
-\newblock // The Encyclopedia of Optimization. Springer, Heidelberg. 2009. P.
+\newblock The direct global optimization algorithm.
+\newblock The Encyclopedia of Optimization. Springer, Heidelberg. 2009. P.
   725--735.
@@
 \newblock Parallel branch and bound for global optimization with combination of
-  lipschitz bounds //
+  lipschitz bounds.
 \newblock Optim. Method. Softw. 1997. Vol. 26, No. 3. P. 487--498.
@@
 {Riquelme} N., {Von Lucken} C., {Baran} B.
-\newblock Performance metrics in multi-objective optimization
-\newblock // 2015 Latin American Computing Conference (CLEI). 2015. P. 1--11.
+\newblock Performance metrics in multi-objective optimization.
+\newblock2015 Latin American Computing Conference (CLEI). 2015. P. 1--11.
 \newblock \doi{10.1109/clei.2015.7360024}
@@
 {Grishagin V.}
-\newblock {Operating characteristics of some global search algorithms}
-\newblock // Problems of Stochastic Search. Zinatne, Riga. 1978. (in Russian)
+\newblock {Operating characteristics of some global search algorithms.}
+\newblock Problems of Stochastic Search. Zinatne, Riga. 1978. (in Russian)
 
@@
 {Norkin V. I.}
-\newblock Towards Pijavskyj's method for solving common global optimization problem //
+\newblock Towards Pijavskyj's method for solving common global optimization problem.
 \newblock J. Comput. Math. Math. Phys. 1992. Vol. 32, no. 7. P. 992--1006. (in Russian)

### 2020-02-27 2019/Automation/Barkalov.tex
all done exept figures

@@
 \usepackage{soul}
+\usepackage{cite}
+\usepackage[labelsep=period]{caption}
+%						 format=hang,
+%            font=onehalfspacing,
+%            margin=\the\parindent,
+%            figurename=Рисунок,
 
@@
 \authors{
-Р.Г.~СТРОНГИН, д-р~физ.-мат.~наук (strongin@unn.ru)\\
-В.П.~ГЕРГЕЛЬ, д-р~техн.~наук (gergel@unn.ru)\\
+Р.Г.~СТРОНГИН, д-р~физ.-мат.~наук (strongin@unn.ru),\\
+В.П.~ГЕРГЕЛЬ, д-р~техн.~наук (gergel@unn.ru),\\
 К.А.~БАРКАЛОВ, канд.~физ.-мат.~наук (barkalov@vmk.unn.ru)\\
 (Нижегородский государственный университет им. Н.И. Лобачевского)}
+
+\KeyWords{пример}
 
@@
 %А нужен ли рисунок?
-Возникающие при этом подзадачи и взаимосвязи между ними отображены на рис. \ref{fig0}.
+Возникающие при этом подзадачи и взаимосвязи между ними отображены на рис. \ref{fig0}. \marginpar{\framebox{Рис. 1}}
 Наглядно видно, что структура взаимосвязей имеет форму дерева,
@@
 причем функции $\varphi_N(y_1,...,y_N)$ на $N$-м уровне являются листьями дерева задач, их значения вычисляются непосредственно.
-
-\begin{figure}
-\begin{center}%[scale=0.3][width=0.75\textwidth]
-  \includegraphics[width=0.7\textwidth]{fig1.pdf} 
-  \caption{Взаимосвязи между подзадачами в многошаговой схеме.}\label{fig0} 
-\end{center}
-\end{figure}
 
@@
 	\item 
-	итерация глобального поиска состоит в выборе одной (наиболее перспективной) задачи из множества имеющихся задач, в которой проводится одно испытание; точка проведения нового испытания определяется в соответствии с алгоритмом из раздела \ref{SectionPeano};
+	итерация глобального поиска состоит в выборе одной (наиболее перспективной) задачи из множества имеющихся задач, в которой проводится одно испытание; точка проведения нового испытания определяется в соответствии с алгоритмом из подраздела \ref{SectionPeano};
 	\item
@@
 Краткое описание основных шагов блочной адаптивной схемы редукции размерности состоит в следующем.
-Пусть вложенные подзадачи вида (\ref{block_nested_i}) решаются с помощью алгоритма глобального поиска, описанного в разделе \ref{SectionPeano}.
+Пусть вложенные подзадачи вида (\ref{block_nested_i}) решаются с помощью алгоритма глобального поиска, описанного в подразделе \ref{SectionPeano}.
 Тогда каждой подзадаче (\ref{block_nested_i}) можно присвоить некоторое числовое значение, называемое характеристикой этой задачи.
@@
 максимальную характеристику интервалов, сформированных в данной задаче.
-В соответствии с правилом вычисления характеристик (\ref{R}), чем выше значение характеристики, тем более перспективной является подзадача для продолжения поиска в ней глобального минимума исходной задачи (\ref{main_problem}).
+В соответствии с правилом вычисления характеристик (\ref{R}) чем выше значение характеристики, тем более перспективной является подзадача для продолжения поиска в ней глобального минимума исходной задачи (\ref{main_problem}).
 Поэтому на каждой итерации выбирается подзадача с максимальной характеристикой для проведения в ней очередного испытания.
@@
 
-Отметим новые элементы предлагаемой схемы редукции размерности.
+\noindent Отметим новые элементы предлагаемой схемы редукции размерности:
 \begin{itemize}
@@
 С другой стороны, он позволяет (при необходимости) увеличивать необходимую точность решения задачи (\ref{main_problem}) без проведения повторного решения всех подзадач (\ref{block_nested_i}).
-Например, начальная стадия оптимизации может быть выполнена с достаточно грубой точностью, что позволит быстро получить некоторую оценку решения; далее (после анализа результатов) точность может быть повышена и процесс глобального поиска может быть продолжен.
+Например, начальная стадия оптимизации может быть выполнена с достаточно грубой точностью, что позволит быстро получить некоторую оценку решения; далее (после анализа результатов) точность может быть повышена и процесс глобального поиска может быть продолжен;
 	\item
-	В исходной многошаговой схемой редукции размерности решение каждой возникающей подзадачи проводится <<до конца>>, т.е.
-порождения новой поздадачи не произойдет, пока не будет решена текущая подзадача.
+	В исходной многошаговой схеме редукции размерности решение каждой возникающей подзадачи проводится ``до конца'', т.е.
+порождения новой подзадачи не произойдет, пока не будет решена текущая подзадача.
 Но точка $\overline{u_i}$ ее решения, вообще говоря, не будет совпадать с $i$-й компонентой вектора решения $y^* = (u_1^\ast,...,u_i^\ast,...,u_M^\ast)$ исходной многомерной задачи.
-В адаптивной схеме подобные бесперспективные подзадачи могут быть отбракованы уже на начальном этапе их решения.
+В адаптивной схеме подобные бесперспективные подзадачи могут быть отбракованы уже на начальном этапе их решения;
 \item
-Наличие правила выбора подзадачи при выполнения очередной итерации в рамках адаптивной схемы позволяет получать различные ее варианты, соответствующие разным правилам выбора подзадачи.
-Например, как исходная многошаговая схема редукции размерности, так и схема редукции размерности с помощью единственной кривой Пеано могут быть получены как частные случаи нового подхода.
+Наличие правила выбора подзадачи при выполнении очередной итерации в рамках адаптивной схемы позволяет получать различные ее варианты, соответствующие разным правилам выбора подзадачи.
+Например, как исходная многошаговая схема редукции размерности, так и схема редукции размерности с помощью единственной кривой Пеано могут быть получены как частные случаи нового подхода;
 \item
@@
 Стандартный подход к сравнению алгоритмов глобальной оптимизации основан на решении указанными методами серии задач, выбранных случайно из некоторого класса.
-Будем использовать два класса многоэкстремальных тестовых функций (GKLS \textit{Simple}, GKLS \textit{Hard}), описанные в \cite{Gaviano03}.
+Будем использовать два класса многоэкстремальных тестовых функций (GKLS \textit{Simple}, GKLS \textit{Hard}), описанных в \cite{Gaviano03}.
 
-В работах \cite{Grishagin18,Sovrasov19} было экспериментально установлено, что алгоритм глобального поиска (АГП) как с использованием разверток, так и в сочетании с адаптивной схемой редукции размерности превосходит многие известные алгоритмы глобальной оптимизации, включая методы DIRECT и DIRECT\textit{l}.
-Поэтому в данном исследовании мы ограничимся сравнением вариантов АГП в сочетании с различными схемами редукции размерности.
+В \cite{Grishagin18,Sovrasov19} было экспериментально установлено, что алгоритм глобального поиска (АГП) как с использованием разверток, так и в сочетании с адаптивной схемой редукции размерности превосходит многие известные алгоритмы глобальной оптимизации, включая методы DIRECT и DIRECT\textit{l}.
+Поэтому в данном исследовании ограничимся сравнением вариантов АГП в сочетании с различными схемами редукции размерности.
 
@@
 Задачу будем считать решенной, если алгоритм сгенерировал точку $y^k$ очередного испытания в окрестности решения $y^\ast$, т.е.
-$\left\|y^k-y^\ast\right\| < \delta \left\|b-a\right\|$, где $\delta = 0.01$, $a$ и $b$ --- границы области поиска $D$.
+$\left\|y^k-y^\ast\right\| < \delta \left\|b-a\right\|$, где $\delta = 10^{-2}$, $a$ и $b$ --- границы области поиска $D$.
 
@@
 \centering
+\captionsetup{labelfont=bf}
 \caption{Среднее число испытаний на классах GKLS \textit{Simple} и \textit{Hard}, $N=2$}\label{tab1}
-\begin{tabular}{lccc}
-\hline\noalign{\smallskip}
+\begin{tabular}{| lcc |}
+\hline 
+%\smallskip
  & GKLS \textit{Simple} &  GKLS \textit{Hard} \\
-\noalign{\smallskip}\hline\noalign{\smallskip}
+%\noalign{\smallskip}
+\hline
+%\noalign{\smallskip}
  $K_e$ &  252 & 674 \\
@@
  $K_a$ &  279 & 815 \\
-\noalign{\smallskip}\hline
+%\noalign{\smallskip}
+\hline
 \end{tabular}
@@
 
-\begin{figure}
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{2DSimple.pdf} \\ (a)}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{2DHard.pdf} \\ (b)}
-\end{minipage}
-\caption{Операционные характеристики на классах GKLS \textit{Simple} (a) и \textit{Hard} (b), $N=2$}
-\label{fig2}
-\end{figure}
-
 Первая серия экспериментов была проведена на двумерных задачах классов GKLS \textit{Simple} и GKLS \textit{Hard} (100 функций каждого класса).
 В табл.~\ref{tab1}  представлено среднее число испытаний, выполненных АГП с использованием разверток ($K_e$), многошаговой схемы ($K_n$) и адаптивной многошаговой схемы ($K_a$).
-На рис.~\ref{fig2}(a,b) приведены операционные характеристики алгоритмов, полученные на указанных классах задач.
-Непрерывная линия соответствует алгоритму с использованием разверток, короткий пунктир --- адаптивной многошаговой схеме, длинный пунктир --- многошаговой схеме.
-Результаты экспериментов показывают, что АГП с использованием адаптивной многошаговой схемы демонстрирует примерно одинаковое быстродействие  по сравнению с развертками, и оба они значительно превосходят алгоритм, использующий многошаговую схему.
-Поэтому в дальнейших экспериментах мы ограничимся сравнением различных вариантов адаптивной схемы редукции размерности.
+На рис.~\ref{fig2} \marginpar{\framebox{Рис.
+2}}
+приведены операционные характеристики алгоритмов, полученные на указанных классах задач.
+Сплошная линия соответствует алгоритму с использованием разверток,точечная линия --- адаптивной многошаговой схеме, штриховая линия --- многошаговой схеме.
+Результаты экспериментов показывают, что АГП с использованием адаптивной многошаговой схемы демонстрирует примерно одинаковое быстродействие  по сравнению с развертками и оба они значительно превосходят алгоритм, использующий многошаговую схему.
+Поэтому в дальнейших экспериментах ограничимся сравнением различных вариантов адаптивной схемы редукции размерности.
+
+\begin{table}
+\captionsetup{labelfont=bf}
+\centering
+\caption{Среднее число испытаний на классах GKLS \textit{Simple} и \textit{Hard}, $N=4$}\label{tab2}
+\begin{tabular}{| l c c |}
+\hline
+%\noalign{\smallskip}
+ &    GKLS \textit{Simple} &  GKLS \textit{Hard} \\
+%\noalign{\smallskip}
+\hline
+%\noalign{\smallskip}
+ $K_a$  &  21 747 & 35 633 \\
+ $K_{ba}$ &  15 942 & 33 206 \\
+%\noalign{\smallskip}
+\hline
+\end{tabular}
+\end{table}
 
@@
 
-На рис.~\ref{fig3}(a,b) приведены операционные характеристики алгоритмов, полученные на классах GKLS \textit{Simple} и GKLS \textit{Hard}.
-Пунктирная линия соответствует алгоритму с использованием адаптивной, а непрерывная ---  блочной адаптивной схемы.
-Результаты экспериментов показывают, что использование блочной адаптивной схемы дает ощутимый выигрыш по числу испытаний (до 35\%) по сравнению с исходной схемой редукции размерности.
-
-\begin{table}
-\centering
-\caption{Среднее число испытаний на классах GKLS \textit{Simple} и \textit{Hard}, $N=4$}\label{tab2}
-\begin{tabular}{lcc}
-\hline\noalign{\smallskip}
- &    GKLS \textit{Simple} &  GKLS \textit{Hard} \\
-\noalign{\smallskip}\hline\noalign{\smallskip}
- $K_a$  &  21 747 & 35 633 \\
- $K_{ba}$ &  15 942 & 33 206 \\
-\noalign{\smallskip}\hline
-\end{tabular}
-\end{table}
-
-\begin{figure}[ht]
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{4DSimple.pdf} \\ (a)}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{4DHard.pdf} \\ (b)}
-\end{minipage}
-\caption{Операционные характеристики на классах GKLS \textit{Simple} (a) и \textit{Hard} (b), $N=4$}
-\label{fig3}
-\end{figure}
+На рис.~\ref{fig3} \marginpar{\framebox{Рис. 3}}
+приведены операционные характеристики алгоритмов, полученные на классах GKLS \textit{Simple} и GKLS \textit{Hard}.
+Точечная линия соответствует алгоритму с использованием адаптивной, а сплошная ---  блочной адаптивной схемы.
+Результаты экспериментов показывают, что использование блочной адаптивной схемы дает ощутимый выигрыш по числу испытаний (до 35 \%) по сравнению с исходной схемой редукции размерности.
 
@@
 Среднее число испытаний, выполненных АГП с использованием разверток, составило 102 987, тогда как с использованием блочной адаптивной схемы  --- 75 390. 
-На рис.~\ref{fig4} приведены операционные характеристики алгоритмов. Пунктирная линия соответствует алгоритму с использованием разверток, а непрерывная --- блочной адаптивной схемы.
-
-\begin{figure}
-\center
-\includegraphics[width=0.5\linewidth]{6DSimple.pdf}
-\caption{Операционные характеристики на классе GKLS \textit{Simple}, $N=6$}
-\label{fig4}
-\end{figure}
+На рис.~\ref{fig4} \marginpar{\framebox{Рис. 4}}
+приведены операционные характеристики алгоритмов. Точечная линия соответствует алгоритму с использованием разверток, а сплошная --- блочной адаптивной схемы.
 
@@
 
-В данной работе предложена обобщенная адаптивная схема редукции размерности для задач глобальной оптимизации, комбинирующая использование кривых Пеано и схему вложенной (рекурсивной) оптимизации.
+В данной статье предложена обобщенная адаптивная схема редукции размерности для задач глобальной оптимизации, комбинирующая использование кривых Пеано и схему вложенной (рекурсивной) оптимизации.
 Для решения редуцированных подзадач используется алгоритм глобального поиска.
@@
 \bibitem{Evtushenko71}
-{\it	Евтушенко Ю.Г.} Численный метод поиска глобального экстремума функций (перебор на неравномерной сетке) // Ж. вычисл. матем. и матем. физ. 1971. T. 11, №6, С. 1390--1403.
+{\it	Евтушенко Ю.Г.} Численный метод поиска глобального экстремума функций (перебор на неравномерной сетке) // Журн. вычисл. матем. и матем. физ. 1971. T.~11. №~6. С.~1390--1403.
 
 \bibitem{Pijavski72}
-{\it	Пиявский С.А.} Один алгоритм отыскания абсолютного экстремума функций // Ж. вычисл. матем. и матем. физ. 1972. Т. 12, № 4. С. 888--896.
+{\it	Пиявский С.А.} Один алгоритм отыскания абсолютного экстремума функций // Журн. вычисл. матем. и матем. физ. 1972. Т. 12. № 4. С. 888--896.
 
 \bibitem{Shubert72}
-{\it Shubert B.O.} A sequential method seeking the global maximum of a function // SIAM J. Numer. Anal. 1972. V. 9. P. 379–388.
+{\it Shubert B.O.} A Sequential Method Seeking the Global Maximum of a Function // SIAM J. Numer. Anal. 1972. V. 9. P. 379–388.
 
 \bibitem{Jones93} 
-{\it Jones D.R., Perttunen C.D., Stuckman B.E.} Lipschitzian optimization without the Lipschitz constant // J. Optim. Theory Appl. 1993. V. 79. No. 1. P. 157--181.
+{\it Jones D.R., Perttunen C.D., Stuckman B.E.} Lipschitzian Optimization without the Lipschitz Constant // J. Optim. Theory Appl. 1993. V. 79. No. 1. P. 157--181.
 
 \bibitem{Pinter96}
-{\it Pinter J. D.} Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications). Dordrecht: Kluwer Academic Publishers, 1996.
+{\it Pinter J.D.} Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications). Dordrecht: Kluwer Acad. Publishers, 1996.
 
 \bibitem{Zilinskas08}
-{\it {\v Z}ilinskas J.} Branch and bound with simplicial partitions for global optimization // Math. Model. Anal. 2008. V. 13. No. 1. P. 145--159.
+{\it {\v Z}ilinskas J.} Branch and Bound with Simplicial Partitions for Global Optimization // Math. Model. Anal. 2008. V. 13. No. 1. P.~145--159.
 
 \bibitem{Evtushenko07}
-{\it Евтушенко Ю.Г., Малкова В.У., Станевичюс А.-И. А.} Распараллеливание процесса поиска глобального экстремума // Автомат. и телемех. 2007. № 5. C. 46--58
+{\it Евтушенко Ю.Г., Малкова В.У., Станевичюс А.-И.А.} Распараллеливание процесса поиска глобального экстремума // АиТ. 2007. №~5. C.~46--58.
+
+{\it Evtushenko Yu.G., Malkova V.U., Stanevichyus A.A.} Parallelization of the Global Extremum Searching Process // Autom. Remote Control. 2007. V. 68. No. 5. P.~787--798.
 
 \bibitem{Evtushenko09}
-{\it Евтушенко Ю.Г., Малкова В.У., Станевичюс А.-И. А.} Параллельный поиск глобального экстремума функций многих переменных // Ж. вычисл. матем. и матем. физ. 2009. Т. 49, №2. C. 255--269.
+{\it Евтушенко Ю.Г., Малкова В.У., Станевичюс А.-И.А.} Параллельный поиск глобального экстремума функций многих переменных // Журн. вычисл. матем. и матем. физ. 2009. Т.~49. №~2. C.~255--269.
 
 \bibitem{Jones09}%Проверить формат ссылки!!!
-{\it Jones D. R.} The DIRECT global optimization algorithm. In: Floudas C. A., Pardalos P. M. (eds.) The Encyclopedia of Optimization, Second Edition.  Springer. 2009. P. 725--735.
+{\it Jones D.R.} The DIRECT global optimization algorithm / Floudas C.A., Pardalos P.M. (eds.) The Encyclopedia of Optimization. Second Edition.  Springer, 2009. P.~725--735.
 
 \bibitem{Zilinskas10}
-{\it Paulavi{\v c}ius R., {\v Z}ilinskas J., Grothey A.}: Investigation of selection strategies in branch and bound algorithm with simplicial partitions and combination of Lipschitz bounds // Optim.
+{\it Paulavi{\v c}ius R., {\v Z}ilinskas J., Grothey A.} Investigation of Selection Strategies in Branch and Bound Algorithm with Simplicial Partitions and Combination of Lipschitz Bounds // Optim.
 Lett.
@@
 V.
-4(1).
+4 (1).
 P.
-173–--83
+173--83.
 
 \bibitem{Evtushenko13}
-{\it Evtushenko Y. G., Posypkin M. A.} A deterministic approach to global box-constrained optimization // Optim. Lett. 2013. V. 7(4). P. 819--829
+{\it Evtushenko Y.G., Posypkin M.A.} A Deterministic Approach to Global Box-Constrained Optimization // Optim. Lett. 2013. V. 7 (4). P. 819--829.
 
 \bibitem{Kvasov13}
-{\it Квасов Д.Е., Сергеев Я.Д.} Методы липшицевой глобальной оптимизации в задачах управления // Автомат. и телемех. 2013. № 9. C. 3--19.
+{\it Квасов Д.Е., Сергеев Я.Д.} Методы липшицевой глобальной оптимизации в задачах управления // АиТ. 2013. № 9. C. 3--19.
+
+{\it Kvasov D.E., Sergeyev Ya.D.} Lipschitz Global Optimization Methods in Control Problems // Autom. Remote Control. 2013. V. 74. No. 9. P. 1435--1448.
 
 \bibitem{Paulavicius16}
-{\it Paulavi{\v c}ius R., {\v Z}ilinskas J.} Advantages of simplicial partitioning for Lipschitz optimization problems with linear constraints // Optim. Lett. 2016. V. 10(2). P. 237--246.
+{\it Paulavi{\v c}ius R., {\v Z}ilinskas J.} Advantages of Simplicial Partitioning for Lipschitz Optimization Problems with Linear Constraints // Optim. Lett. 2016. V.~10~(2). P.~237--246.
 
 \bibitem{Kureichik04}
-{\it Гладков Л.А., Курейчик В.В., Курейчик В.М.} Генетические алгоритмы. Учебное пособие. М.: ФИЗМАТЛИТ, 2004.
+{\it Гладков Л.А., Курейчик В.В., Курейчик В.М.} Генетические алгоритмы. Уч. пос. М.: Физматлит, 2004.
 
 \bibitem{Karpenko08}
-{\it Карпенко А.П.} Современные алгоритмы поисковой оптимизации. Алгоритмы, вдохновленные природой: учебное пособие. М.: Издательство МГТУ им. Н. Э. Баумана, 2014.
+{\it Карпенко А.П.} Современные алгоритмы поисковой оптимизации. Алгоритмы, вдохновленные природой. Уч. пос. М.: Изд-во МГТУ им. Н.Э. Баумана, 2014.
 
 \bibitem{Kvasov18}
-{\it Kvasov D.E., Mukhametzhanov M.S.} Metaheuristic vs. deterministic global optimization algorithms: The univariate case // Appl. Math. Comput. 2018. V. 318. P. 245--259.
+{\it Kvasov D.E., Mukhametzhanov M.S.} Metaheuristic vs. Deterministic Global Optimization Algorithms. The Univariate Case // Appl. Math. Comput. 2018. V.~318. P.~245--259.
 
 \bibitem{Sergeyev18}
-{\it Sergeyev Y., Kvasov D., Mukhametzhanov M.} On the efficiency of nature-inspired metaheuristics in expensive global optimization with limited budget // Sci.
-Rep. V. 8(1). Art. No. 435.
+{\it Sergeyev Y., Kvasov D., Mukhametzhanov M.} On the Efficiency of Nature-Inspired Metaheuristics in Expensive Global Optimization with Limited Budget // Sci.
+Rep. V.~8~(1). Art. No. 435.
 
 \bibitem{Neymark66a}
-{\it Неймарк Ю.И., Стронгин Р.Г.} Поиск экстремума функций по принципу максимума информации // Автомат. и телемех. 1966. № 11. С. 113--118.
+{\it Неймарк Ю.И., Стронгин Р.Г.} Поиск экстремума функций по принципу максимума информации // АиТ. 1966. №~11. С.~113--118.
+
+{\it Neimark Yu.I., Strongin R.G.} Function Extremum Search with the Use of Information Maximum Principle // Autom. Remote Control. 1966. V. 27. No. 1. P.~101--105.
 
 \bibitem{Neymark66b}
-{\it Неймарк Ю.И., Стронгин Р.Г.} Информационный подход к задаче поиска экстремума функций // Изв. АН СССР, Техническая кибернетика. 1966. № 1. С. 17--26.
+{\it Неймарк Ю.И., Стронгин Р.Г.} Информационный подход к задаче поиска экстремума функций // Изв. АН СССР, Технич. кибернетика. 1966. №~1. С.~17--26.
 
 \bibitem{Strongin70}
-{\it Стронгин Р.Г.} Многоэкстремальная минимизация // Автомат. и телемех. 1970. № 7. С. 63--67.
+{\it Стронгин Р.Г.} Многоэкстремальная минимизация // АиТ. 1970. №~7. С.~63--67.
+
+{\it Strongin R.G.} Multiextremal Minimization // Autom. Remote Control. 1970. V. 7. P.~1085--1088.
 
@@
 \bibitem{Strongin86}
-{\it Стронгин Р.Г., Маркин Д.Л.} Минимизация многоэкстремальных функций при невыпуклых ограничениях // Кибернетика. 1986. №4. С.63--69.
+{\it Стронгин Р.Г., Маркин Д.Л.} Минимизация многоэкстремальных функций при невыпуклых ограничениях // Кибернетика. 1986. №~4. С.~63--69.
 
 \bibitem{Strongin87}
-{\it Маркин Д.Л., Стронгин Р.Г.} Метод решения многоэкстремальных задач с невыпуклыми ограничениями, использующий априорную информацию об оценках оптимума // Ж.
+{\it Маркин Д.Л., Стронгин Р.Г.} Метод решения многоэкстремальных задач с невыпуклыми ограничениями, использующий априорную информацию об оценках оптимума // Журн.
 вычисл.
@@
 1987.
-Т.27, № 1.
-С.52--61.
+Т.~27.
+№~1.
+С.~52--61.
 
 \bibitem{Strongin93}
-{\it Маркин Д.Л., Стронгин Р.Г.} О равномерной оценке множества слабоэффективных точек в многоэкстремальных многокритериальных задачах оптимизации // Ж.
+{\it Маркин Д.Л., Стронгин Р.Г.} О равномерной оценке множества слабоэффективных точек в многоэкстремальных многокритериальных задачах оптимизации // Журн.
 вычисл.
@@
 физ.
-1993 Т.
-33, № 2.
-С.
-195--205.
+1993.
+Т.~33.
+№~2.
+С.~195--205.
 
 \bibitem{Strongin13}
-{\it Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013.
+{\it Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Изд-во МГУ, 2013.
 
 \bibitem{Strongin91}
-{\it Стронгин Р.Г.} Параллельная многоэкстремальная оптимизация с использованием множества разверток // Ж. вычисл. матем. и матем. физ. 1991. T. 31, № 8. С. 1173--1185.
+{\it Стронгин Р.Г.} Параллельная многоэкстремальная оптимизация с использованием множества разверток // Журн. вычисл. матем. и матем. физ. 1991. T.~31. №~8. С.~1173--1185.
 
 \bibitem{Strongin00}
-{\it Strongin R.G., Sergeev Ya.D.} Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Dordrecht: Kluwer Academic Publishers, 2000.
+{\it Strongin R.G., Sergeev Ya.D.} Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Dordrecht: Kluwer Acad. Publishers, 2000.
 
 \bibitem{Grishagin97}
-{\it Grishagin V.A., Sergeyev Y.D., Strongin R.G.} Parallel characteristical algorithms for solving problems of global optimization // J. Glob. Optim. 1997. V. 10(2). P. 185--206.
+{\it Grishagin V.A., Sergeyev Y.D., Strongin R.G.} Parallel Characteristical Algorithms for Solving Problems of Global Optimization // J. Glob. Optim. 1997. V.~10~(2). P. 185--206.
 
 \bibitem{Grishagin01}
-{\it Sergeyev Y., Grishagin V.} Parallel asynchronous global search and the nested optimization scheme // J. Comput. Anal. Appl. 2001. V. 3(2). P. 123--145.
+{\it Sergeyev Y., Grishagin V.} Parallel Asynchronous Global Search and the Nested Optimization Scheme // J. Comput. Anal. Appl. 2001. V.~3~(2). P.~123--145.
 
 \bibitem{Grishagin16}
-{\it Gergel V., Grishagin V., Gergel A.} Adaptive nested optimization scheme for multidimensional global search // J. Glob. Optim. 2016. V. 66(1). P. 35--51.
+{\it Gergel V., Grishagin V., Gergel A.} Adaptive Nested Optimization Scheme for Multidimensional Global Search // J. Glob. Optim. 2016. V.~66~(1). P.~35--51.
 
 \bibitem{Barkalov2014}
-{\it Barkalov K., Gergel V.} Multilevel scheme of dimensionality reduction for parallel
-global search algorithms // In: OPT-i 2014 -- 1st International Conference on Engineering and Applied Sciences Optimization, Proceedings. 2014. P. 2111--2124.
+{\it Barkalov K., Gergel V.} Multilevel Scheme of Dimensionality Reduction for Parallel Global Search Algorithms / OPT-i 2014 --- 1st Int.
+Conf.
+on Engineering and Applied Sciences Optimization.
+Proc.
+2014.
+P.
+2111--2124.
 
 \bibitem{Grishagin2001}
-{\it Sergeyev Y., Grishagin V.} Parallel asynchronous global search and the nested optimization scheme // J. Comput. Anal. Appl. 2001. V. 3(2). P. 123--145.
+{\it Sergeyev Y., Grishagin V.} Parallel Asynchronous Global Search and the Nested Optimization Scheme // J. Comput. Anal. Appl. 2001. V. 3~(2). P. 123--145.
 
 \bibitem{Grishagin2015}
-{\it Gergel V., Grishagin V., Israfilov R.} Local tuning in nested scheme of global optimization // Procedia Computer Science. 2015. V. 51(1). P. 865--874
+{\it Gergel V., Grishagin V., Israfilov R.} Local Tuning in Nested Scheme of Global Optimization // Procedia Computer Sci. 2015. V.~51~(1). P.~865--874.
 
 \bibitem{Barkalov2016}
-{\it Barkalov K., Lebedev I.} Solving multidimensional global optimization problems using graphics accelerators // CCIS. 2016. V. 687. P. 224--235.
+{\it Barkalov K., Lebedev I.} Solving Multidimensional Global Optimization Problems using Graphics Accelerators // CCIS. 2016. V. 687. P. 224--235.
 
 \bibitem{Grishagin2016_1}
-{\it Grishagin V., Israfilov R., Sergeyev Y.}: Comparative efficiency of dimensionality reduction schemes in global optimization // AIP Conference Proceedings. 2016. V. 1776, art. no. 060011.
+{\it Grishagin V., Israfilov R., Sergeyev Y.} Comparative Efficiency of Dimensionality Reduction Schemes in Global Optimization // AIP Conf. Proc. 2016. V. 1776. Art. No. 060011.
 
 \bibitem{Grishagin18}
-{\it Grishagin V., Israfiov R., Sergeyev Y.} Convergence conditions and numerical comparison of global optimization methods based on dimensionality reduction schemes // Appl.
+{\it Grishagin V., Israfiov R., Sergeyev Y.} Convergence Conditions and Numerical Comparison of Global Optimization Methods Based on Dimensionality Reduction Schemes // Appl.
 Math.
@@
 \bibitem{Gergel19}
-{\it Gergel V., Grishagin V., Israfilov R.} Parallel dimensionality reduction for multiextremal optimization problems // LNCS. 2019. V. 11657. P. 166--178.
+{\it Gergel V., Grishagin V., Israfilov R.} Parallel Dimensionality Reduction for Multiextremal Optimization Problems // LNCS. 2019. V. 11657. P. 166--178.
 
 \bibitem{Gaviano03}
-{\it Gaviano M., Lera D., Kvasov D.E., Sergeyev Ya.D.} Software for generation of classes of test functions with known local and global minima for global optimization // ACM Trans.
+{\it Gaviano M., Lera D., Kvasov D.E., Sergeyev Ya.D.} Software for Generation of Classes of Test Functions with Known Local and Global Minima for Global Optimization // ACM Trans.
 Math.
@@
 \bibitem{Sovrasov19}
-{\it Sovrasov V.} Comparison of several stochastic and deterministic derivative-free global optimization algorithms // LNCS. 2019. V. 11548. P. 70--81.
-
-%\bibitem{Sergeyev00} Kvasov, D.E., Pizzuti, C., Sergeyev, Y.D.: Local tuning and partition strategies for diagonal GO methods. Numer. Math. 94(1), 93--106 (2003)
-%\bibitem{Sergeyev06} Sergeyev, Ya.D., Kvasov, D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim. 16(3), 910--937 (2006)
-
-%\bibitem{third} {\it Третий Т.Т.} Публикация в трудах конференции // Тр. Ин-та такого-то РАН. 2000. Т. 1. № 2. С. 3--4.
-%\bibitem{forth} {\it Четвертый Ч.Ч.} Публикация по теме в серийном издании или сборнике / Сб. научн. тр. к 20-летию Института. Новосибирск, Изд-во <<Пирогов>>, 2000. Т. 1. № 2. С. 3--4.
-%\bibitem{fifth} {\it Fifth F.} Some journal publication in English // Appl. Math. Comput. J., Elsevier Publ. 1925. V. 501. No. 1. P. 1234--5678.
-%\bibitem{sixth} {\it Sixth J.Th.} A book in English. Boston: Springer, 1991.
+{\it Sovrasov V.} Comparison of Several Stochastic and Deterministic Derivative-Free Global Optimization Algorithms // LNCS. 2019. V. 11548. P. 70--81.
 
@@
 
-\AdditionalInformation{Стронгин Р.Г.}{Нижегородский государственный университет им. Н.И. Лобачевского, президент, Нижний Новгород}{strongin@unn.ru}
+\clearpage
 
-\AdditionalInformation{Гергель В.П.}{Нижегородский государственный университет им.
-Н.И.
-Лобачевского, директор института информационных технологий, математики и механики, Нижний Новгород}{gergel@unn.ru}
+\begin{center}
+\bfseries ?enoiee e noaoua Aaeaiaei 
+\end{center}
 
-\AdditionalInformation{Баркалов К.А.}{Нижегородский государственный университет им.
-Н.И.
-Лобачевского, доцент кафедры математического обеспечения и суперкомпьютерных технологий института информационных технологий, математики и механики, Нижний Новгород}{konstantin.barkalov@itmm.unn.ru}
+\vfill
+
+\begin{figure}[h]
+\begin{center}%[scale=0.3][width=0.75\textwidth]
+  \includegraphics[width=0.7\textwidth]{fig1.pdf} 
+  %\caption*{Взаимосвязи между подзадачами в многошаговой схеме.}\label{fig0} 
+  \caption{}\label{fig0} 
+\end{center}
+\end{figure}
+
+%\begin{figure}[h]
+  %\centering
+  %\includegraphics[scale=1.0]{pic01}
+  %\caption*{Aaeaiaei, ?en. 1}
+%\end{figure}
+\vfill
+
+\clearpage
+
+\begin{figure}[h]
+\begin{minipage}{0.5\linewidth}
+ \center{а}  \\ \center{\includegraphics[width=1.0\linewidth]{2DSimple.pdf}}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+ \center{б} \\ \center{\includegraphics[width=1.0\linewidth]{2DHard.pdf}}
+\end{minipage}
+%\caption{Операционные характеристики на классах GKLS \textit{Simple} (а) и \textit{Hard} (б), $N=2$.}
+\caption{} \label{fig2}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[h]
+\begin{minipage}{0.5\linewidth}
+\center{а} \\ \center{\includegraphics[width=1.0\linewidth]{4DSimple.pdf}}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+\center{б} \\ \center{\includegraphics[width=1.0\linewidth]{4DHard.pdf}}
+\end{minipage}
+%\caption{Операционные характеристики на классах GKLS \textit{Simple} (а) и \textit{Hard} (б), $N=4$.}
+\caption{} \label{fig3}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+\center
+\includegraphics[width=0.5\linewidth]{6DSimple.pdf}
+%\caption{Операционные характеристики на классе GKLS \textit{Simple}, $N=6$.}
+\caption{} \label{fig4}
+\end{figure}
+
+\clearpage
+
+\begin{center}
+\bfseries Iiaiene e ?enoieai noaoue Aaeaiaei 
+\end{center}
+
+\begin{itemize}%[font=\normalfont]
+  \item [Рис. 1.] Noaiaoe?aneia ecia?a?aiea iaaieoiiai iiaaana
+\end{itemize}
 

### 2020-02-27 2019/Automation/Barkalov.tex
Figures placed on separate pages

@@
 \begin{center}
-\bfseries ?enoiee e noaoua Aaeaiaei 
+\bfseries Рисунки к статье  
 \end{center}
@@
 
-%\begin{figure}[h]
-  %\centering
-  %\includegraphics[scale=1.0]{pic01}
-  %\caption*{Aaeaiaei, ?en. 1}
-%\end{figure}
 \vfill
@@
 \clearpage
+
+\phantom{f}
+\vfill
 
@@
 
+\vfill
+
 \clearpage
+
+\phantom{f}
+\vfill
 
@@
 \end{figure}
+
+\vfill
 
@@
 \begin{center}
-\bfseries Iiaiene e ?enoieai noaoue Aaeaiaei 
+\bfseries Подписи к рисункам статьи  
 \end{center}
@@
 \begin{itemize}%[font=\normalfont]
-  \item [Рис. 1.] Noaiaoe?aneia ecia?a?aiea iaaieoiiai iiaaana
+  \item [Рис. 1.] Взаимосвязи между подзадачами в многошаговой схеме.
+  \item [Рис. 2.] Операционные характеристики на классах GKLS \textit{Simple} (а) и \textit{Hard} (б), $N=2$.
+  \item [Рис. 3.] Операционные характеристики на классах GKLS \textit{Simple} (а) и \textit{Hard} (б), $N=4$.
+	\item [Рис. 4.] Операционные характеристики на классе GKLS \textit{Simple}, $N=6$.
 \end{itemize}

### 2020-02-28 2019/Automation/Barkalov.tex
All Correct

@@
 
-\KeyWords{пример}
+\KeyWords{глобальная оптимизация, многоэкстремальные функции, редукция размерности, кривые Пеано, рекурсивная оптимизация. }
 
@@
 
-\noindent Отметим новые элементы предлагаемой схемы редукции размерности:
+Отметим новые элементы предлагаемой схемы редукции размерности:
 \begin{itemize}
@@
 приведены операционные характеристики алгоритмов, полученные на указанных классах задач.
-Сплошная линия соответствует алгоритму с использованием разверток,точечная линия --- адаптивной многошаговой схеме, штриховая линия --- многошаговой схеме.
+Сплошная линия соответствует алгоритму с использованием разверток, точечная линия --- адаптивной многошаговой схеме, штриховая линия --- многошаговой схеме.
 Результаты экспериментов показывают, что АГП с использованием адаптивной многошаговой схемы демонстрирует примерно одинаковое быстродействие  по сравнению с развертками и оба они значительно превосходят алгоритм, использующий многошаговую схему.
@@
 
-\begin{center}
-\bfseries Рисунки к статье  
-\end{center}
+%\begin{center}
+%\bfseries Рисунки к статье  
+%\end{center}
+\phantom{f}
 
@@
 \begin{center}
-\bfseries Подписи к рисункам статьи  
+\bfseries Список подрисуночных подписей  
 \end{center}

### 2020-03-12 2020/PAVT_Sovrasov/slides/slides.tex
PAVT: add a draft of slides

--- 2019-03-31 2019/PAVT/slides/slides.tex
@@
 
+\usepackage[T2A]{fontenc}
+\usepackage[utf8]{inputenc}
+\inputencoding{utf8}
+\usepackage[russian]{babel}
 \usepackage{listings}
@@
 
-\setromanfont{CMU Serif}
-%\setsansfont{CMU Sans Serif}
-\setmathfont{Latin Modern Math}
+\setmainfont{CMU Sans Serif}
+\setromanfont{CMU Sans Serif}
+\setsansfont{CMU Sans Serif}
+
 \setlength{\tabcolsep}{1pt}
@@
 \usepackage{polyglossia}
+\setmainlanguage{russian}
 %\setbeamertemplate{itemize item}{\color{black}$\blacktriangleright$}
@@
 
-\title{Comparison Of Several Sequential And Parallel Derivative-free Global Optimization Algorithms}
-\author{\underline{\textbf{Vladislav~Sovrasov}} and \textbf{Semen~Bevzuk}}
-\institute{Lobachevsky State University of Nizhni Novgorod}
+\title{Параллельный алгоритм для получения равномерного приближения решений множества задач глобальной оптимизации с нелинейными ограничениями}
+\author{\underline{\textbf{В.В.~Соврасов}} \and \textbf{К.А.~Баркалов}}
+\institute{Нижегородский государственный университет им. Н.И. Лобачевского}
 \date{}
@@
 \begin{frame}
-  \frametitle{Problem statement}
+  \frametitle{Постановка задачи}
   \begin{columns}
@@
 \begin{frame}
-  \frametitle{Methods}
+  \frametitle{Conclusions}
     \begin{itemize}
-    \item[$\square$] \textbf{Deteriminstic}
-    \begin{itemize}
-      \item Have complicated internal structure in multidimensional case;
-      \item Usually store and use the whole history of trials accumulated during search;
-      \item Require non-trivial parallelization schemes;
-      \item Under some assumptions convergence to the global solution is guaranteed;
+      \item
     \end{itemize}
-    \item[$\square$] \textbf{Stochastic}
-    \begin{itemize}
-      \item Have relative simple internal structure;
-      \item Require constant amount of memory to store internal state of some random process or individuals of population;
-      \item In most cases allows wide-scale parallelization by trials or by running many copies of the same method
-      with different seeds in parallel.
-      \item Convergence is guaranteed in probabilistic sense only.
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Methods}
-  In this work considered the following solvers available in open-source:
-  \begin{itemize}
-    \item[$\square$] \textbf{Deteriminstic}
-    \begin{itemize}
-      \item DIRECT, DIRECT$l$;
-      \item AGS, AGS$l$;
-    \end{itemize}
-    \item[$\square$] \textbf{Stochastic}
-    \begin{itemize}
-      \item Multi Level Single Linkage;
-      \item StoGO;
-      \item Differential Evolution;
-      \item Controlled Random Search;
-      \item Dual Simulated Annealing;
-      \item Ant colony optimization;
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Test problems}
-  \begin{columns}
-    \begin{column}{0.5\textwidth}
-      Generator GKLS was employed to construct 8 sets of 100 test problems:
-      \begin{displaymath}
-        \begin{matrix}
-          f(x)=
-          \left\{
-          \begin{matrix}
-          C_i(x), x \in S_i, i\in 2,\dots ,m \\
-          \Vert x-T \Vert^2 + t, x\not\in S_2,\dots,S_m
-          \end{matrix} \right.
-        \end{matrix}
-      \end{displaymath}
-
-      The generator allows to adjust:
-      \begin{itemize}
-        \item the number of local minimas;
-        \item the size of the global minima attraction region;
-        \item the space dimension.
-      \end{itemize}
-    \end{column}
-    \begin{column}{0.5\textwidth}
-      \centerline{\includegraphics[width=0.9\textwidth]{gkls_color.png}}
-    \end{column}
-  \end{columns}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results of sequential methods}
-
-  \begin{table}
-  \begin{center}
-  \caption{Averaged number of trials executed by sequential methods for solving the test
-  optimization problems}
-  \resizebox{\textwidth}{!}{%
-    \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
-      \hline
-      & AGS & AGS\(l\) & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
-    \hline
-    \(F_{GR}\)     & 193.1 &  \textbf{158.3} & 400.3 & 182.3 & 214.9 & 947.2 & 691.2 & 1257.3 &
-  1336.8 \\
-    \hline
-    GKLS 2d Simple &  254.9 & 217.6 & 510.6 & \textbf{189.0} & 255.2 & 556.8 & 356.3 & 952.2
-  & 1251.5 \\
-    \hline
-    GKLS 2d Hard   &  728.7 & \textbf{488.0} & 844.7 & 985.4 & 1126.7 & 1042.5 & 1637.9 &
-  1041.1 & 2532.2 \\
-    \hline
-    GKLS 3d Simple &  1372.1 & 1195.3 & 4145.8 & \textbf{973.6} & 1477.8 & 4609.2 & 2706.5 &
-  5956.94 & 3856.1 \\
-    \hline
-    GKLS 3d Hard   &  3636.1 & \textbf{1930.5} & 6787.0 & 2298.7 & 3553.3 & 5640.1 & 4708.4 &
-  6914.3 & 7843.2 \\
-    \hline
-    GKLS 4d Simple &  26654.1 & 11095.7 & 37436.8 & \textbf{7824.3} & 15994.1 & 41514.3 &
-  21417.9 & 19157.7 & 59895.4 \\
-    \hline
-    GKLS 4d Hard   &  54536.8 &  \textbf{23167.8} &  73779.3 &  23204.4 &  54489.9 &  80247.2
-  &  68815.5 &  27466.1 &  109328.1  \\
-    \hline
-    GKLS 5d Simple &  29810.0 & 11529.0 & 143575.0 & \textbf{7166.5} & 13970.5 & 52647.6 &
-  34255.3 & 73074.5 & 91580.4 \\
-    \hline
-    GKLS 5d Hard   &  113129.1 & 67652.7 & 165192.8 & \textbf{66327.4} & 164390.6 & 138766.2
-  & 116973.1 & 105496.9 & 155123.8 \\
-    \hline
-  \end{tabular}}
-    \label{tab:trials}
-  \end{center}
-  \end{table}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results of sequential methods}
-  \begin{table}
-  \begin{center}
-  \caption{Number of test optimization problems solved by sequential methods}
-    \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
-      \hline
-      & AGS & AGS\(l\) & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
-    \hline
-    \(F_{GR}\)     &  100 & 100 & 76  & 100 & 100 & 97  & 96  & 96  & 67\\
-    \hline
-    GKLS 2d Simple &  100 & 100 & 85  & 100 & 100 & 100 & 100 & 98  & 90\\
-    \hline
-    GKLS 2d Hard   &  100 & 100 & 74  & 100 & 100 & 100 & 93  & 85  & 77 \\
-    \hline
-    GKLS 3d Simple &  100 & 97  & 75  & 100 & 100 & 100 & 89  & 86  & 44 \\
-    \hline
-    GKLS 3d Hard   &  100  & 99   & 72   & 100  & 99   & 100  & 88   & 77   & 43 \\
-    \hline
-    GKLS 4d Simple &  100 & 100 & 46  & 100 & 100 & 94  & 78  & 59  & 16 \\
-    \hline
-    GKLS 4d Hard   &  100 & 100 & 47  & 99  & 97  & 94  & 72  & 32  & 10  \\
-    \hline
-    GKLS 5d Simple &  100 & 100 & 68  & 100 & 100 & 98  & 100 & 77  & 9  \\
-    \hline
-    GKLS 5d Hard   &  97  & 99  & 42  & 100 & 90  & 79  & 84  & 48  & 8 \\
-    \hline
-    \end{tabular}
-    \label{tab:solved}
-  \end{center}
-  \end{table}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Basic optimization method}
-  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
-  \begin{enumerate}
-    \setlength{\itemindent}{.1in}
-    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
-    \item[Step 2.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
-    \item[Step 3.] Choose \(p\) intervals \((x_{t_j-1}, x_{t_j})\) with the greatest characteristics and
-    compute objective \(\varphi(y(x^{k+j}))\) in points chosen using the decision rule \(d\):
-    \begin{displaymath}
-      x^{k+1+j}=d(t)\in (x_{t_j-1}, x_{t_j}),\:j=\overline{1,p}
-    \end{displaymath}
-    \item[Step 4.] If \(x_{t_j}-x_{t_j-1}<\varepsilon\) for one of \(j=\overline{1,p}\), stop the method.
-  \end{enumerate}
-  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
-\end{frame}
-
-\begin{frame}
-  \begin{center}
-  \frametitle{Dimension reduction}
-  Peano-type curve \(y(x)\) allows to reduce the dimension of the original problem:
-  \begin{gather}
-    \lbrace y\in \mathbb{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\} \nonumber \\
-    \min\{\varphi(y): y\in D\}=\min\{\varphi(y(x)): x\in [0,1]\} \nonumber
-  \end{gather}
-  \(y(x)\) is non-smooth function which continuously maps the segment \([0,1]\) to the hypercube \(D\).
-  \begin{figure}[ht]
-    \vspace*{-0.5cm}
-    \subfloat{{\includegraphics[width=.35\textwidth]{peano3d.png} }}
-    \subfloat{{\includegraphics[width=.35\textwidth]{peano2d.png} }}
-  \end{figure}
-\end{center}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Parallel Globalizer solver}
-  \begin{figure}[ht]
-    \includegraphics[width=0.95\textwidth]{evolvents_parallel.pdf}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Parallel MIDACO solver}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Cluster environment}
-  \begin{center}
-  The computational experiments have been carried out on the Lobachevsky supercomputer at  State University of Nizhni Novgorod.
-  One node includes 2 Intel Sandy Bridge E5-2660 2.2 GHz processors and 64 GB RAM.
-
-  Each node runs under CentOS 7 Linux with GCC 4.8 compiler and Intel MPI library.
-\end{center}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results of applying the parallel algorithm}
-  \begin{table}
-    \centering
-    \caption{Averaged numbers of iterations executed by the parallel algorithms for solving the test
-  optimization problems}
-    \label{tab:iterations}
-    \begin{tabular}{cccccccc}
-      \cline{3-8}\noalign{\smallskip}
-      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{Globalizer} & &
-  \multicolumn{2}{c}{MIDACO}   \\
-      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
-      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
-  \textit{Hard}  \\
-      \noalign{\smallskip}\hline
-      I & \textbf{1 cluster node}
-        & \textit{1} &   25270 (100)  & 55180 (99) & & 27645 (98) & 72068 (71)  \\
-      &  & \textit{16} & 1765 (100)  & 3714 (100) & &  1640 (97) & 4304 (70) \\
-      \hline \noalign{\smallskip}
-  II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
-    & \textit{1} & 13056 (100) & 22938 (99) & & 10558 (89) & 27128 (73) \\
-  &   & \textit{16} & 732 (100) & 1759 (100)  &  & 1130 (92) & 2254 (73) \\
-      \hline \noalign{\smallskip}
-  III & \textbf{4 cluster nodes} %\multirow{3}{*}{}
-    & \textit{1}  & 5016 (100) & 12703 (100) & & 5777 (94) & 15980 (75) \\
-  & & \textit{16} & 367 (100) & 776 (100) & & 529 (88) &  1264 (66) \\
-      \hline \noalign{\smallskip}
-  VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-    & \textit{1}  & 2103 (100) & 5063 (100) & & 2847 (97) & 9853 (83)\\
-  & & \textit{16} & 145 (100)  & 310 (100)  & & 272 (83) & 774 (57)\\
-      \hline \noalign{\smallskip}
-  V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
-    & \textit{1}  & 1155 (100) & 2399 (100) & & 2233 (98) & 6022 (86) \\
-  & & \textit{16} & 76 (100)  & 159 (100)  & & 168 (87) & 393 (53)\\
-      \noalign{\smallskip}\hline
-    \end{tabular}
-  \end{table}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results of applying the parallel algorithm}
-  \begin{table}
-    \centering
-    \caption{Speedup of parallel computations executed by the parallel algorithms}
-    \label{tab:speedup}
-    \begin{tabular}{cccccccc}
-      \cline{3-8}\noalign{\smallskip}
-      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{Globalizer} & &
-  \multicolumn{2}{c}{MIDACO}   \\
-      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
-      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
-  \textit{Hard}  \\
-      \noalign{\smallskip}\hline
-  I  & \textbf{1 cluster node}  %\multirow{3}{*}{}
-      & \textit{1}   & 25270 (27.3s) & 55180 (61.8s) & & 27645 (32.2s) & 72068 (132.5s)  \\
-    &  & \textit{16} & 14.3(11.7) & 14.9(12.6)  & &  16.9(14.4) & 16.7(14.4) \\
-    \hline \noalign{\smallskip}
-  II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
-    & \textit{1}      &   1.9(1.9) & 2.4(2.2)  & & 2.6(1.7) & 2.7(2.3) \\
-    &   & \textit{16} & 34.5(20.4) & 31.4(18.8) & & 24.5(18.8) & 32.0(29.1) \\
-      \noalign{\smallskip}\hline	\noalign{\smallskip}
-  III  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
-  & \textit{1}      & 5.0(4.6) & 4.3(4.1) & & 4.8(3.9) & 4.5(4.4) \\
-  &   & \textit{16} & 68.8(23.7) & 71.2(25.8) & & 52.3(32.9) & 57.0(50.2) \\
-    \noalign{\smallskip}\hline	\noalign{\smallskip}
-  VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
-  & \textit{1}    & 12.0(8.7) & 10.9(8.3) & & 9.7(8.2)    & 7.3(9.0)    \\
-  & & \textit{16} & 174.1(4.7) & 177.8(5.3) & & 101.6(51.6) & 93.1(84.1) \\
-      \noalign{\smallskip}\hline
-      V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
-      & \textit{1}    & 21.9(11.4)  & 23.0(12.8)  & & 12.4(11.1)  & 12.0(14.8)  \\
-      & & \textit{16} & 333.5(2.7) & 347.9(3.0) & & 164.5(82.0) & 183.2(129.7) \\
-          \noalign{\smallskip}\hline
-    \end{tabular}
-  \end{table}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results of applying the parallel algorithm}
-  \begin{figure}[ht]
-    \centering
-    \subfloat[$p=1$]{{\includegraphics[width=.5\textwidth]{speedup_midaco_globalizer_p_1.pdf}}\label{fig:speedup_p1}}
-    \subfloat[$p=16$]{{\includegraphics[width=.5\textwidth]{speedup_midaco_globalizer_p_16.pdf}}\label{fig:speedup_p16}}
-    \caption{Speedup demonstrated by the parallel algorithms when solving problems from the GKLS 4d Hard class}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results of applying the parallel algorithm}
-  \begin{figure}[ht]
-    \centering
-    \subfloat[$p=1$]{{\includegraphics[width=.5\textwidth]{time_midaco_globalizer_p_1.pdf}}\label{fig:time_p1}}
-    \subfloat[$p=16$]{{\includegraphics[width=.5\textwidth]{time_midaco_globalizer_p_16.pdf}}\label{fig:time_p16}}
-    \caption{Averaged execution time of the parallel algorithms when solving problems from the GKLS 4d Hard class}
-  \end{figure}
-\end{frame}
-
-%\begin{frame}
-%  \frametitle{Results of applying the parallel algorithm}
-%\end{frame}
-
-\begin{frame}
-  \frametitle{Conclusions}
-    %Already done:
-    \begin{itemize}
-      \item 1
-    \end{itemize}
-    %Future work:
-    %\begin{itemize}
-    %  \item
-    %\end{itemize}
 \end{frame}

### 2020-03-13 2019/NUMTA_Barkalov/SoftComputing/article.tex
correct some typos

@@
 \usepackage{graphicx}
-\usepackage{lineno,hyperref,amsmath}
+\usepackage{hyperref,amsmath}
+\usepackage[pagewise]{lineno}
 \usepackage{multirow}
@@
 \begin{document}
+
+%\linenumbers
 
@@
 
-	Since we have assumed that $x'$ is a limit point for the IM-DL and  $x' \neq \bar x$, then there exists such a sequence of the intervals $[x_{t-1}, x_t], t=t(k)$ that $x' \in [x_{t-1}, x_t], R_{loc}(t)>0$, and $R_{loc}(t) \rightarrow 0$.
+	Since we have assumed that $x'$ is a limit point for the IA-DL and  $x' \neq \bar x$, then there exists such a sequence of the intervals $[x_{t-1}, x_t], t=t(k)$ that $x' \in [x_{t-1}, x_t], R_{loc}(t)>0$, and $R_{loc}(t) \rightarrow 0$.
 This property of the characteristics of the intervals shrinking to the limit point has been established when proving the theorem on the convergence of the IA in \cite{Strongin2000}.
 
-	Therefore, for large enough, the following inequality will be true
+	Therefore, for large enough $k$, the following inequality will be true
 \begin{equation}

### 2020-03-16 2019/NUMTA_Barkalov/SoftComputing/article.tex
Aswers for reviewer

@@
 \usepackage[misc]{ifsym}
+\usepackage{xcolor}
 % etc.
@@
 	\textbf{Proof}. Let us consider two possible cases of the values of the indices $\nu(x_{i-1})$ and $\nu(x_i)$.
-\begin{enumerate} 
-  \item If $\nu(x_{i-1}) \neq \nu(x_i)$, then according to  formulae (\ref{R_2}), (\ref{R_3}) 
+
+  1. If $\nu(x_{i-1}) \neq \nu(x_i)$, then according to  formulae (\ref{R_2}), (\ref{R_3}) 
 \begin{equation}\label{R_glob_1}
@@
 \end{equation}
-  \item If $\nu(x_{i-1}) = \nu(x_i) = M$, then, taking into account (\ref{current_lower_bounds}), the following inequality will hold
+
+{ \color{red}
+	2. If $\nu(x_{i-1}) = \nu(x_i) = M$, then, taking into account (\ref{current_lower_bounds}), the following inequality will be valid
 $$
-	z_i+z_{i-1}-2z_M^* = |z_i-z_{i-1}|\leq \mu_M\Delta_i .
+	z_i+z_{i-1}-2z_M^* = |z_i-z_{i-1}| = \mu \Delta_i \leq \mu_M\Delta_i,
 $$
-\end{enumerate} 
+where $\mu = |z_i-z_{i-1}|/\Delta_i$.
+
 From here, it will follow that for any value $r_M>1$ the following inequalities will hold
 \begin{align}
-	R(i) & = \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{z_i+z_{i-1}-2z_M^*}{r_M\mu_M} \geq \nonumber \\
-	& \geq  \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\mu_M\Delta_i}{r_M\mu_M} \geq \nonumber \\
-  & \geq \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\Delta_i}{r_M} = \nonumber \\
-	& = \Delta_i + \Delta_i\frac{(z_i-z_{i-1})^2/\Delta_i^2}{r_M^2\mu_M^2\Delta_i^2} - 2\frac{\Delta_i}{r_M} \geq \nonumber \\
-	& \geq \Delta_i + \frac{\Delta_i}{r_M^2} - 2\frac{\Delta_i}{r_M} = \Delta_i \left( 1-\frac{1}{r_M} \right)^2>0.
+	R(i) & = \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{z_i+z_{i-1}-2z_M^*}{r_M\mu_M} = \nonumber \\
+	& = \Delta_i + \frac{\mu^2\Delta_i}{r_M^2\mu_M^2} - 2\frac{\mu\Delta_i}{r_M\mu_M} = \nonumber \\
+  & = \Delta_i\left(1 + \left(\frac{\mu}{r_M\mu_M}\right)^2 - 2\frac{\mu}{r_M\mu_M}\right) = \nonumber \\
+	& = \Delta_i\left(1 - \frac{\mu}{r_M\mu_M}\right)^2 \geq  \Delta_i \left( 1-\frac{1}{r_M}\right)^2 > 0.
 \end{align}
-And because $r_{M}^{glob} > r_{M}^{loc} > 1$, then
+
+And because $r_{M}^{glob} > r_{M}^{loc} > 1$ and $\mu \leq \mu_M$, then
 \begin{align}\label{R_glob_greater_R_loc}
-	R_{glob}(i) & = \Delta_i\left(1-\frac{1}{r_M^{glob}} \right)^2 > \nonumber \\
-	& > \Delta_i\left(1-\frac{1}{r_M^{loc}} \right)^2 = R_{loc}(i).
+	R_{glob}(i) &= \Delta_i\left(1-\frac{\mu}{r_M^{glob}\mu_M} \right)^2 > \nonumber \\
+							&> \Delta_i\left(1-\frac{\mu}{r_M^{loc}\mu_M} \right)^2 = R_{loc}(i).
 \end{align}
+
+}	
+%\begin{enumerate} 
+  %\item If $\nu(x_{i-1}) \neq \nu(x_i)$, then according to  formulae (\ref{R_2}), (\ref{R_3}) 
+%\begin{equation}\label{R_glob_1}
+	%R(i)=R_{glob}(i)=R_{loc}(i)=2\Delta_i>0. 
+%\end{equation}
+  %\item If $\nu(x_{i-1}) = \nu(x_i) = M$, then, taking into account (\ref{current_lower_bounds}), the following inequality will hold
+%$$
+	%z_i+z_{i-1}-2z_M^* = |z_i-z_{i-1}|\leq \mu_M\Delta_i .
+%$$
+%\end{enumerate} 
+%From here, it will follow that for any value $r_M>1$ the following inequalities will hold
+%\begin{align}
+	%R(i) & = \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{z_i+z_{i-1}-2z_M^*}{r_M\mu_M} \geq \nonumber \\
+	%& \geq  \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\mu_M\Delta_i}{r_M\mu_M} \geq \nonumber \\
+  %& \geq \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{\Delta_i}{r_M} = \nonumber \\
+	%& = \Delta_i + \Delta_i\frac{(z_i-z_{i-1})^2/\Delta_i^2}{r_M^2\mu_M^2\Delta_i^2} - 2\frac{\Delta_i}{r_M} \geq \nonumber \\
+	%& \geq \Delta_i + \frac{\Delta_i}{r_M^2} - 2\frac{\Delta_i}{r_M} = \Delta_i \left( 1-\frac{1}{r_M} \right)^2>0.
+%\end{align}
+%And because $r_{M}^{glob} > r_{M}^{loc} > 1$, then
+%\begin{align}\label{R_glob_greater_R_loc}
+	%R_{glob}(i) & = \Delta_i\left(1-\frac{1}{r_M^{glob}} \right)^2 > \nonumber \\
+	%& > \Delta_i\left(1-\frac{1}{r_M^{loc}} \right)^2 = R_{loc}(i).
+%\end{align}
+
 The statement is proved.
@@
 	& + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} = \\
-	& = \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
+	& = { \color{red} \Delta_i\left(1-\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} } > 0.
 \end{align*}
@@
 
-	Since we have assumed that $x'$ is a limit point for the IA-DL and  $x' \neq \bar x$, then there exists such a sequence of the intervals $[x_{t-1}, x_t], t=t(k)$ that $x' \in [x_{t-1}, x_t], R_{loc}(t)>0$, and $R_{loc}(t) \rightarrow 0$.
+	%	Since we have assumed that $x'$ is a limit point for the IA-DL and  $x' \neq \bar x$, then there exists such a sequence of the intervals $[x_{t-1}, x_t], t=t(k)$ that $x' \in [x_{t-1}, x_t], R_{loc}(t)>0$, and $R_{loc}(t) \rightarrow 0$.
 This property of the characteristics of the intervals shrinking to the limit point has been established when proving the theorem on the convergence of the IA in \cite{Strongin2000}.
+
+{ \color{red}	
+	At the beginning of the proof of the theorem we have assumed that $x'$ is a limit point for the IA-DL and  $x' \neq \bar x$.
+This assumption implies that in the method IA-DL in the formula (\ref{R_max}) for computing the characteristic, the second component takes the highest values, i.e.
+$R_{glob}(t) < \rho R_{loc}(t)$.
+This fact corresponds to using the original index algorithm (IA) with parameters $r_\nu = r_\nu^{loc}>1, \; 1 \leq \nu \leq m+1$.
+
+In \cite{Strongin2000} it has been proven that for any limit point $x'$ of the sequence $\{x^k\}$, generated by IA while solving the problem at any $r_\nu > 1, \; 1 \leq \nu \leq m+1$, there exists a sequence
+of the intervals $[x_{t-1}, x_t], \; t=t(k)$, contracting to the point $x'$, i.e.
+$x' \in [x_{t-1}, x_t], \; t=t(k)$.
+Characteristics $R(t)$ of the intervals of the specified sequence will be positive and $R(t) \rightarrow 0$ at $k \rightarrow \infty$.
+}
 
@@
 \begin{equation}
-	R_{glob}(j) > \rho R_{loc}(t)
+	R_{glob}(j) > \rho R_{loc}(t).
 \end{equation}
-and (according to Rules 5 and 6 of the IA-DL method) the next iteration falls into the interval $[x_{j-1}, x_j]$, which contradicts our assumption regarding the non-limit nature of the point $\bar x$.
+{\color{red}
+
+Indeed, both $j$ and $t$ depend from $k$, as the numbering of the trial points (by the lower index) changes at every iteration and the numbering of the intervals changes respectively.
+However, number $j(k)$ corresponds to the same interval with fixed characteristic \linebreak $R_{glob}(j(k))>0$, whereas number $t(k)$ corresponds to different intervals with different characteristics \linebreak $R_{loc}(t(k))>0$, where $R_{loc}(t(k)) \rightarrow 0$.
+From this we can conclude that the condition $R_{glob}(j) > \rho R_{loc}(t)$ is fulfilled for a large enough value of $k$.
+}
+
+Hence, according to Rules 5 and 6 of the IA-DL method the next iteration falls into the interval $[x_{j-1}, x_j]$, which contradicts our assumption regarding the non-limit nature of the point $\bar x$.
 The theorem is proven.

### 2020-03-29 2020/PAVT_Sovrasov/slides/slides.tex
PAVT: update vm version

@@
       \end{displaymath}
-      \(\varphi(y)\) is multiextremal objective function, which satisfies the Lipschitz condition:
+      \(\varphi(y)\) -- многоэкстремальная функция, удовлятворяющая условию Липшица:
       \begin{displaymath}
@@
       \end{displaymath}
-      where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes \(l_2\) norm in \(\mathbb{R}^N\)
-      space.
+      где \(L>0\) константа Липшица и \(||\cdot||\) обозначает \(l_2\) норму в пространстве \(\mathbb{R}^N\).
     \end{column}
@@
 \begin{frame}
-  \frametitle{Conclusions}
+  \frametitle{Заключение}
     \begin{itemize}

### 2020-03-29 2020/PAVT_Sovrasov/vm_journal/vm_paper.tex
PAVT: update vm version

@@
 \usepackage{doi}
-\usepackage{subfig}
 
@@
 \title{ПАРАЛЛЕЛЬНЫЙ АЛГОРИТМ ДЛЯ ПОЛУЧЕНИЯ РАВНОМЕРНОГО ПРИБЛИЖЕНИЯ
-РЕШЕНИЙ МНОЖЕСТВА ЗАДАЧ ГЛОБАЛЬНОЙ ОПТИМИЗАЦИИ С НЕЛИНЕЙНЫМИ ОГРАНИЧЕНИЯМИ
+РЕШЕНИЙ МНОЖЕСТВА ЗАДАЧ ГЛОБАЛЬНОЙ ОПТИМИЗАЦИИ С~НЕЛИНЕЙНЫМИ ОГРАНИЧЕНИЯМИ%
 \footnote{Статья рекомендована к публикации программным комитетом Международной
@@
 
-\author{\copyright~2020~г. В.В.~Соврасов, К.A.~Баркалов}
+\author{\copyright~2020 В.В.~Соврасов, К.A.~Баркалов}
 
@@
 экстремума было оценено примерно с одинаковым качеством. Обычно серию из \(q\) задач решают либо последовательно, либо
-параллельно порциями по \(p,\:p<<q\) задач, где \(p\) --- количество параллельных вычислительных устройств.
+параллельно порциями по \(p\ll q\) задач, где \(p\) --- количество параллельных вычислительных устройств.
 Такой подход ведет к тому, что в каждый момент времени до окончания вычислений
@@
 Эти генераторы порождают задачи без нелинейных ограничений, поэтому в дополнение к ним использована
-система GCGen \footnote{Исходный код системы доступен по ссылке https://github.com/UNN-ITMM-Software/GCGen} \cite{GergelBarkalov2019}, позволяющая генерировать задачи с ограничениями на основе произвольных нелинейных
+система GCGen\footnote{Исходный код системы доступен по ссылке https://github.com/UNN-ITMM-Software/GCGen} \cite{GergelBarkalov2019}, позволяющая генерировать задачи с ограничениями на основе произвольных нелинейных
 функций.
@@
 был сгенерирован смешанный класс, состоящий из 50 задач с двухмерными функциями GKLS и 50 задач с функциями \(F_{GR}\).
-На рис. \ref{fig:isolines} представлены примеры линий уровня рассматриваемых задач. Допустимая область закрашена.
+На \figref{pic/2.png} и \figref{pic/4.png} представлены примеры линий уровня рассматриваемых задач. Допустимая область закрашена.
 
-\begin{figure}[ht]
-    \centering
-    \subfloat[Решение внутри допустимой области]{
-    {\includegraphics[width=.5\textwidth]{pic/2.png}}}
-    \subfloat[Решение на границе допустимой области]{
-    {\includegraphics[width=.5\textwidth]{pic/4.png}}}
-    \caption{Линии уровня и точки испытаний ИАГП в двух синтетических задачах}
-    \label{fig:isolines}
-\end{figure}
+\fig{scale=0.3}{pic/2.png}{Пример линий уровня задачи с решением внутри допустимой области}
+\fig{scale=0.3}{pic/4.png}{Пример линий уровня задачи с решением на границе допустимой области}
 
@@
 Результаты решения тестовых задач последовательной и параллельной версией модифицированного ИАГП
-для решения множества задач представлены в таблице \ref{tab:speedup}. Для всех двухмерных классов задач параметр \(r=4.7\).
+для решения множества задач представлены в \tabref{tab:speedup}. Для всех двухмерных классов задач параметр \(r=4.7\).
 В случае трехмерных задач \(r=4.7,\: \varepsilon_\nu=0.1\).
@@
 
-\begin{table}
-  \centering
-  \caption{Результаты экспериментов на наборах синтетических задач}
-  \label{tab:speedup}
+\tab{tab:speedup}{Результаты экспериментов на наборах синтетических задач}{
   \begin{tabular}{c|c|cccc}
@@
   \end{tabular}
-\end{table}
+}
 
 Для того, чтобы показать равномерную сходимость все тестовые задачи были также решены ИАГП
-в режиме решения отдельных задач. На рис. \ref{fig:devs_mixed} указаны графики величин
+в режиме решения отдельных задач. На \figref{pic/mixed_2d_max.pdf} и \figref{pic/mixed_2d_avg.pdf} указаны графики величин
 средних и максимальных расстояний от реальных оптимумов до их текущих оценок при решении
@@
 
-\begin{figure}[ht]
-    \centering
-    \subfloat[\(D_{max}\)]{
-    \label{fig:max_dev} {\includegraphics[width=.5\textwidth]{pic/mixed_2d_max.pdf}}}
-    \subfloat[\(D_{avg}\)]{
-    \label{fig:avg_dev} {\includegraphics[width=.5\textwidth]{pic/mixed_2d_avg.pdf}}}
-    \caption{Динамика величин \(D_{avg}\) и \(D_{max}\) в процессе решения множества двухмерных задач,
+\fig{scale=0.5}{pic/mixed_2d_max.pdf}{Динамика величины \(D_{max}\) в процессе решения множества двухмерных задач,
 порождённых двумя разными генераторами GKLS и \(F_{GR}\)}
-    \label{fig:devs_mixed}
-\end{figure}
+\fig{scale=0.5}{pic/mixed_2d_avg.pdf}{Динамика величины \(D_{avg}\) в процессе решения множества двухмерных задач,
+порождённых двумя разными генераторами GKLS и \(F_{GR}\)}
 
@@
 
-\begin{figure}[ht]
-    \centering
-    \subfloat[ИАГП, раздельное решение задач]{
-    \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{pic/single_mco.pdf}}}
-    \subfloat[ИАГП для множества задач]{
-    \label{fig:mco_pareto_2} {\includegraphics[width=.5\textwidth]{pic/multi_mco.pdf}}}
-    \caption{Численные оценки множества Парето в задаче (\ref{eq:mco_probem}), полученные после 2500 испытаний}
-    \label{fig:mco_pareto}
-\end{figure}
+\fig{scale=0.5}{pic/single_mco.pdf}{Численная оценки множества Парето в задаче (\ref{eq:mco_probem}),
+                                    полученная при раздельном решении задач}
+\fig{scale=0.5}{pic/multi_mco.pdf}{Численная оценки множества Парето в задаче (\ref{eq:mco_probem}),
+                                   полученная при совместном решении задач}
 
-На рис. \ref{fig:mco_pareto} представлены графики решений, полученных каждым из методов.
+На \figref{pic/single_mco.pdf} и \figref{pic/multi_mco.pdf}
+представлены графики решений, полученных каждым из методов.
 Все графики качественно совпадают с указанным в \cite{BinhKorn1999} (авторы не предоставили другой
-информации и решениях для сравнения). Видно, что на рис. \ref{fig:mco_pareto_1}
+информации и решениях для сравнения). Видно, что на \figref{pic/single_mco.pdf}
 кривая Парето имеет вогнутости, что не соответствует решению, указанному в \cite{BinhKorn1999} и
@@
 
-%\newpage
-
-{\phantom{eee} \hrule \vskip 4 pt}
+\newpage
 \classify{} % Код MSC не нужен
@@
 
-\title{\uppercase{Parallel global optimization algorithm for obtaining uniform convergence
-when simultaneously solving a set of global optimization problems}}
+\title{\uppercase{Parallel global optimization algorithm for~obtaining uniform convergence
+when~simultaneously solving a~set of~global optimization problems}}
 

### 2020-03-29 2019/NUMTA_Barkalov/SoftComputing/article.tex
fix some reviewer comments

@@
 
-	\textbf{Statement}.
+	\textbf{Lemma 1}.
 Let us assume that the current minimum value $z_M^\ast$  from (\ref{z_estimates}) at the $k^{th}$ iteration is achieved at the boundary point of some interval $[x_{i-1}, x_i], i=i(k)$.
@@
 
-The statement is proved.
+The lemma is proved.
 
@@
 	Let us formulate and prove the convergence conditions for the new algorithm IA-DL.\\
+	{ \color{red}
+
+	\textbf{Lemma 2}. 
+Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and the conditions 1--3 of Theorem \ref{theorem:1} are fulfilled.
+Let $\bar{y}=y(\bar{x})$ be a limit point of the trial sequence $\{y^k\} = \{y(x^k)\}$ generated by the IA for the problem (\ref{problem}) with the parameters $r_{\nu}>1, 1 \leq \nu \leq m+1$.
+Then there will be such a sequence of intervals (corresponding to iterations with numbers $q^i, \; i=1,2,...$)
+\begin{equation}\label{intervals_t}
+\{[x_{t-1},x_t]\}, \; t=t(q^i), \; i=1,2,...,
+\end{equation}
+satisfying the conditions
+\begin{align}
+&\bar{x} \in \bigcap_{i=1}^\infty [x_{t-1},x_t],\label{nested_int} \\ 
+&\lim_{i\rightarrow\infty}\Delta_t = 0,\label{lim_delta_t} \\ 
+&R(t(q^i))>0, \; i=1,2,...,\label{R_g_zero} \\ 
+&\lim_{i\rightarrow\infty} R(t(q^i)) = 0, \label{lim_R}  
+\end{align}
+where $\Delta_t=(x_t-x_{t-1})^{1/N}$,  $R(t)$ and $t$ are from (\ref{R_1})--(\ref{R_3}) and (\ref{MaxR}) respectively.
+
+\textbf{Proof}. As long as $\bar{x}$ is a limit point, there should exist a sequence of trial numbers $q^1, q^2,...$ meeting the conditions
+\[
+\bar{x} \in [x_{t-1}, x_t], \; x^{k+1} \in [x_{t-1}, x_t], \; t=t(k), k=q^1, q^2, ...,
+\]
+meaning that the interval $[x_{t-1}, x_t]$ containing the limit point $\bar{x}$ at the step $k$ is hit by the trial at the subsequent step $k+1$.
+
+From Step 3 and Step 7 of the index algorithm follows that
+\begin{equation}\label{max_gamma}
+\max \{x_t-x^{k+1},x^{k+1}-x_{t-1}\} \leq \gamma (x_t-x_{t-1}),
+\end{equation}
+where $\gamma = \frac{r+1}{2r}<1$  and $ r=\min\{r_\nu:1\leq\nu\leq m+1\}>1$.
+The contraction property (\ref{lim_delta_t}) and the property (\ref{nested_int}) for the product of all intervals (\ref{intervals_t}) are the consequences of (\ref{max_gamma}).
+
+As shown in Lemma 1, at each iteration there exists an interval with strictly positive characteristic. Due to the rule (\ref{MaxR}) this leads to the conclusion that
+intervals from (\ref{intervals_t}) should also have positive characteristics because they are hit by subsequent trials; i.e., the conditions (\ref{R_g_zero}) are necessary.
+
+For intervals with indexes $\nu(x_{i-1}) = \nu(x_{i}) = \nu$ from the second condition of Theorem \ref{theorem:1} and (\ref{K_leq_L}) follows that
+\[
+\frac{\left|z_i-z_{i-1}\right|}{\Delta_i}\leq 2L_\nu\sqrt{N+3},
+\]
+and in accordance with (\ref{z_estimates}) the inequalities 
+\[
+z_i = g_\nu(y(x_i))\geq z_\nu^*, \; \nu=\nu(x_i)
+\]
+holds. Then from the rules (\ref{R_1})--(\ref{R_3}) it follows that $R(i) \leq 2\Delta_i$, if $\nu(x_{i-1}) \neq \nu(x_{i})$, and
+\[
+R(i) \leq \Delta_i\left(1+\left(\frac{2L_\nu\sqrt{N+3}}{r_\nu\mu_\nu}\right)^2\right), 
+\]
+if $\nu(x_{i-1}) = \nu(x_{i}) = \nu$. These inequalities combined with (\ref{lim_delta_t}) lead us to the validity of (\ref{lim_R}).
+
+}
+
 \begin{theorem}\label{theorem:2} Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem \ref{theorem:1} are fulfilled.
@@
 	At every iteration of the IA-DL method there will exist such interval $[x_{j-1}, x_j]$, $j=j(k)$, that $\bar x \in [x_{j-1}, x_j]$.
-According to our assumption that $\bar y = y(\bar x)$ is not the limit point, there will exist such number $k$, starting from which the trial points will not fall into the interval $[x_{j-1}, x_j]$.
-That is, to conduct a new trial at the point $x^{k+1} \in [x_{t-1}, x_t]$ at every iteration such interval $[x_{t-1}, x_t], t=t(k)$ will be selected that $t \neq j$.
+According to our assumption that $\bar y = y(\bar x)$ is not the limit point, there will exist such number $\bar k$, starting from which the trial points will not fall into the interval $[x_{j-1}, x_j]$.
+That is, to conduct a new trial at the point $x^{k+1} \in [x_{t-1}, x_t]$ for each $k\geq \bar k$ such interval $[x_{t-1}, x_t], t=t(k)$ will be selected that $t(k) \neq j(k)$.
 
@@
 
-In \cite{Strongin2000} it has been proven that for any limit point $x'$ of the sequence $\{x^k\}$, generated by IA while solving the problem at any $r_\nu > 1, \; 1 \leq \nu \leq m+1$, there exists a sequence
-of the intervals $[x_{t-1}, x_t], \; t=t(k)$, contracting to the point $x'$, i.e.
-$x' \in [x_{t-1}, x_t], \; t=t(k)$.
+In Lemma 2 it has been proven that for any limit point $x'$ of the sequence $\{x^k\}$, generated by IA while solving the problem at any $r_\nu > 1, \; 1 \leq \nu \leq m+1$, there exists a sequence
+of the intervals $[x_{t-1}, x_t], \; t=t(k)$, contracting to the point $x'$.
 Characteristics $R(t)$ of the intervals of the specified sequence will be positive and $R(t) \rightarrow 0$ at $k \rightarrow \infty$.
@@
 Indeed, both $j$ and $t$ depend from $k$, as the numbering of the trial points (by the lower index) changes at every iteration and the numbering of the intervals changes respectively.
-However, number $j(k)$ corresponds to the same interval with fixed characteristic \linebreak $R_{glob}(j(k))>0$, whereas number $t(k)$ corresponds to different intervals with different characteristics \linebreak $R_{loc}(t(k))>0$, where $R_{loc}(t(k)) \rightarrow 0$.
+%new text begin
+However, $R_{glob}(j(k))>0$ is fixed for each $k \geq \bar k$ because $j(k)$ corresponds to the same interval for $k \geq \bar k$, whereas number $t(k)$ corresponds to different intervals with different characteristics $R_{loc}(t(k)) \rightarrow 0$.
 From this we can conclude that the condition $R_{glob}(j) > \rho R_{loc}(t)$ is fulfilled for a large enough value of $k$.
+%new text end
 }

### 2020-03-30 2020/PAVT_Sovrasov/vm_journal/vm_paper.tex
PAVT: fix some references
PAVT: update intro

@@
 
+Статья имеет следующую структуру. В разделе \ref{sec:problem} рассмотрена
+постановка решаемой задачи, далее в разделе \ref{sec:method} приведено описание
+параллельного метода оптимизациии. Результаты численных экспериментов, подтверждающие
+эффективность рассматриваемого метода приводятся в разделе \ref{sec:exps}.
+В заключении приводится краткая сводка результатов, полученных в ходе работы, и
+указаны направления дальнейших усилий по улучшению программной реализации рассмотренного метода.
+
 \section{Постановка задачи глобальной оптимизации}
+\label{sec:problem}
 В рамках данной работы будем рассматривать следующую постановку задачи глобальной
@@
 \section{Описание метода глобальной оптимизации}
+\label{sec:method}
 
@@
 \section{Результаты численных экспериментов}
+\label{sec:exps}
 
@@
     technique with uniform convergence //
-  \newblock J. of Global Optimization. 2018. Vol. 71, No. 1. P. 21--36.
+  \newblock Journal of Global Optimization. 2018. Vol. 71, no. 1. P. 21--36.
   \newblock \doi{10.1007/s10898-017-0555-4}
@@
   \newblock Best practices for comparing optimization algorithms //
-  \newblock Optimization and Engineering. 2017. Vol. 18, No. 4. P. 815--848.
+  \newblock Optimization and Engineering. 2017. Vol. 18, no. 4. P. 815--848.
   \newblock \doi{10.1007/s11081-017-9366-1}
@@
   \newblock A deterministic approach to global box-constrained optimization //
-  \newblock Optim. Lett. 2013. Vol. 7 P. 819--829.
+  \newblock Optimization Letters. 2013. Vol. 7 P. 819--829.
   \newblock \doi{10.1007/s11590-012-0452-1}
@@
     and global minima for global optimization //
-  \newblock ACM Transactions on Mathematical Software. 2003. Vol. 29, No. 4. P.
+  \newblock ACM Transactions on Mathematical Software. 2003. Vol. 29, no. 4. P.
     469--480.
@@
   \newblock A flexible generator of constrained global optimization test
-    problems. 2019.
+    problems // Proceedings LeGO - 14th International Global Optimization Workshop. 8–21 September 2018, Leiden, The Netherlands. Vol. 2070, no. 1. P. 020009.
   \newblock \doi{10.1063/1.5089976}
@@
     lipschitz bounds //
-  \newblock Optim. Method. Softw.. 1997. Vol. 26, No. 3. P. 487--498.
+  \newblock Optimization Methods and Software. 1997. Vol. 26, no. 3. P. 487--498.
   \newblock \doi{10.1080/10556788.2010.551537}
@@
   \bibitem{SergeyevKvasov2017}
-  Sergeyev Y., Kvasov D. Deterministic Global Optimization. 2017.
+  Sergeyev Y., Kvasov D. Deterministic Global Optimization. 2017. 136 p.
   \newblock \doi{10.1007/978-1-4939-7199-2}
@@
     constraints. Sequential and parallel algorithms.
-  \newblock Kluwer Academic Publishers, Dordrecht. 2000.
+  \newblock Kluwer Academic Publishers, Dordrecht. 2000. 704 p.
   \newblock \doi{10.1007/978-1-4615-4677-1}
@@
   To T.B., Korn B.
-  \newblock MOBES: A multiobjective evolution strategy for constrained
-    optimization problems. 1999.
+  \newblock MOBES: A multiobjective evolution strategy for constrained optimization problems
+  // Proceedings of the third international conference on genetic algorithms (Mendel 97). 1997. P. 176--182.
 
@@
   \newblock // Проблемы статистической
-    оптимизации. Зинатне, Рига. 1978.
+    оптимизации. Зинатне, Рига. 1978. №7. С. 198–206.
 
@@
     задачи глобальной оптимизации //
-  \newblock Ж. вычисл. матем. и матем. физ. 1992. № 32 С. 992--1006.
+  \newblock Журнал вычислительной математики и математической физики. 1992. № 32 С. 992--1006.
 \end{biblio}
@@
   technique with uniform convergence.
-\newblock J. of Global Optimization. 2018. Vol. 71, No. 1. P. 21--36.
+\newblock Journal of Global Optimization. 2018. Vol. 71, No. 1. P. 21--36.
 \newblock \doi{10.1007/s10898-017-0555-4}
@@
 \newblock A deterministic approach to global box-constrained optimization.
-\newblock Optim. Lett. 2013. Vol. 7 P. 819--829.
+\newblock Optimization Letters 2013. Vol. 7 P. 819--829.
 \newblock \doi{10.1007/s11590-012-0452-1}
@@
 \newblock A flexible generator of constrained global optimization test
-  problems. 2019.
+  problems // Proceedings LeGO - 14th International Global Optimization Workshop. 8–21 September 2018, Leiden, The Netherlands. Vol. 2070, no. 1. P. 020009.
 \newblock \doi{10.1063/1.5089976}
@@
   lipschitz bounds.
-\newblock Optim. Method. Softw. 1997. Vol. 26, No. 3. P. 487--498.
+\newblock Optimization Methods and Software. 1997. Vol. 26, No. 3. P. 487--498.
 \newblock \doi{10.1080/10556788.2010.551537}
@@
 \bibitem{SergeyevKvasov2017}
-Sergeyev Y., Kvasov D. Deterministic Global Optimization. 2017.
+Sergeyev Y., Kvasov D. Deterministic Global Optimization. 2017. 136 p.
 \newblock \doi{10.1007/978-1-4939-7199-2}
@@
   constraints. Sequential and parallel algorithms.
-\newblock Kluwer Academic Publishers, Dordrecht. 2000.
+\newblock Kluwer Academic Publishers, Dordrecht. 2000. 704 p.
 \newblock \doi{10.1007/978-1-4615-4677-1}
@@
 To T.B., Korn B.
-\newblock MOBES: A multiobjective evolution strategy for constrained
-  optimization problems. 1999.
+\newblock MOBES: A multiobjective evolution strategy for constrained optimization problems //
+Proceedings of the third international conference on genetic algorithms (Mendel 97). 1997. P. 176--182.
 
@@
 \newblock {Operating characteristics of some global search algorithms.}
-\newblock Problems of Stochastic Search. Zinatne, Riga. 1978. (in Russian)
+\newblock Problems of Stochastic Search. Zinatne, Riga. 1978. No. 7. P. 198–206. (in Russian)
 
@@
 \newblock Towards Pijavskyj's method for solving common global optimization problem.
-\newblock J. Comput. Math. Math. Phys. 1992. Vol. 32, no. 7. P. 992--1006. (in Russian)
+\newblock Computational Mathematics and Mathematical Physics. 1992. Vol. 32, No. 7. P. 992--1006. (in Russian)
 

### 2020-03-30 2020/PAVT_Sovrasov/vm_journal/vm_paper.tex
PAVT: changes in abstrats

@@
 в процессе работы алгоритма погрешности численного решения во всех задачах убывают примерно с одинаковой скоростью.
-Подобные серии задач возникают, если задача глобальной оптимизации имеет дискретный параметр или при решении задачи многокритериальной оптимизации методом свертки критериев.
+Алгоритм присваивает приоритет каждой задаче и на каждой итерации производит вычисления целевых функций в нескольких задачах параллельно.
+При окончании работы метода в произвольный момент времени во всех задачах из решаемой серии будут получены решения сходного качества.
+Серии из нескольких задач возникают, если задача глобальной оптимизации имеет дискретный параметр или, например, при решении задачи многокритериальной оптимизации методом свертки критериев.
 Рассматриваемый алгоритм использует отображения типа кривой Пеано для редукции многомерных задач оптимизации к одномерным.
-Эффективность реализованного алгоритма протестирована на наборах искусственно сгенерированных задач глобальной оптимизации.
+Эффективность реализованного алгоритма протестирована на наборах искусственно сгенерированных задач глобальной оптимизации, а также при решении серии задач, полученной в результате скаляризации задачи многокритериальной оптимизации.
+Также экспериментально подтверждена равномерная сходимость метода.
 
@@
 \maketitle{}
+
 \begin{abstract}%
-  In this work building of a parallel version of a method simultaneously solving a set of
-  constrained global optimization problems is considered. This method converges uniformly
-  to solutions of all the problems. That allows the method to arrange computational resources in an
-  optimal way, since uniform convergence guarantees approximately equal precision
-  of numerical solutions at the whole set of problems at the each iteration of optimization.
-  Sets of similar global optimization problems appear for an instance after scalarization
-  of multi-objective problems or when a global optimization problem has a discrete parameter
-  which takes a finite number of possible values. The considered method uses Peano-type curves
-  to transform multidimensional problems into univariate ones. Efficiency of the implemented
-  parallel algorithm is evaluated on several sets of synthetically generated constrained global optimization problems.
+
+In this work building of a parallel version of a method simultaneously solving a set of constrained global optimization problems is considered.
+This method converges uniformly to solutions of all the problems.
+That allows the method to arrange computational resources in an optimal way, since uniform convergence guarantees approximately equal precision of numerical solutions at the whole set of problems at the each iteration of optimization.
+The algorithm assigns a priority to each problem, and then at the each iteration carries out calculation of objective functions and constraints in several problems in parallel.
+If the method stops at any arbitrary moment, in all the problems numerical sulutions with the similar accuracy will be obtained.
+Sets of similar global optimization problems appear for an instance after scalarization of multi-objective problems or when a global optimization problem has a discrete parameter which takes a finite number of possible values.
+The considered method uses Peano-type curves to transform multidimensional problems into univariate ones.
+Efficiency of the implemented parallel algorithm is evaluated on several sets of synthetically generated constrained global optimization problems and on a scalarized multi-objective problem.
+Also the uniform convergence was confirmed numerically by validation quality of intermediate solutions during the optimization process.
 

### 2020-03-31 2019/NUMTA_Barkalov/SoftComputing/article.tex
Fix second questions of reviewer

@@
 
-	This statement can be interpreted as the incomparability of the local and global characteristics: the global characteristic of the best interval is always larger.
-At the same time, it follows from the formulae (\ref{R_glob_1}), (\ref{R_glob_greater_R_loc}) that
+{ \color{blue}
+
+	This statement can be interpreted as the incomparability of the local and global characteristics: the global characteristic of the best interval is larger if $\nu(x_{i-1})=\nu(x_i)=M$.
+At the same time, it follows from the formula (\ref{R_glob_greater_R_loc}) that
+
 \begin{equation}\label{R_glob_equal_R_loc}
-	R_{glob}(i) = \rho R_{loc}(i),
+	R_{glob}(i) = \left( \frac{1-\frac{\mu}{\mu_M r_M^{glob}}}{1-\frac{\mu}{\mu_M r_M^{loc}}} \right)^2 R_{loc}(i),
 \end{equation}
-where $\rho=\Big(\frac{1-1/r_{M}^{glob}}{1-1/r_{M}^{loc}}\Big)^2$, if $\nu(x_{i-1})=\nu(x_i)=M$, and $\rho=1$, if $\nu(x_{i-1}) \neq \nu(x_i)$.
-In the case of the multiplication by the coefficient $\rho$, the local and global characteristics of the search intervals become comparable, and one can choose the best (i.e.
-the largest) one.
+wherein the maximum value of the equalization coefficient is 
+\begin{equation}\label{new_rho}
+\rho=\left(\frac{1-1/r_{M}^{glob}}{1-1/r_{M}^{loc}}\right)^2 
+\end{equation}
+in the case when $\mu = \mu_M$. %, where $\mu = |z_i-z_{i-1}|/\Delta_i$ and $\mu_M$ is from (\ref{current_lower_bounds}).
+The use of the equalization coefficient equal to $\rho$ from (\ref{new_rho}) in all cases will correspond to equalizing the largest possible difference between the values of $R_{loc}(i)$ and $R_{glob}(i)$ and will contribute to a more local nature of the search when the difference between $R_{loc}(i)$ and $R_{glob}(i)$ is not the largest, i.e.
+when $\mu < \mu_M$.
 
-	Based on the relation (\ref{R_glob_equal_R_loc}), one can formulate an \textit{index algorithm with dual estimate of the Lipschitz constants} (hereafter referred to as IA-DL).
+%Использование нормирующего коэффициента равного $\rho$ из (\ref{new_rho})  будет соответствовать выравниванию наибольшей возможной разницы между значениями $R_{loc}(i)$ и $R_{glob}(i)$ и будет способствовать более ``локальному'' характеру поиска в случае когда разница между $R_{loc}(i)$ и $R_{glob}(i)$ не является максимальной (т.е.
+когда $\mu_M > \mu$).
+
+Based on the relations (\ref{R_glob_equal_R_loc}) and (\ref{new_rho}), one can formulate an \textit{index algorithm with dual estimate of the Lipschitz constants} (hereafter referred to as IA-DL).
+
+}
 
@@
 	Let us formulate and prove the convergence conditions for the new algorithm IA-DL.\\
-	{ \color{red}
+	{ \color{blue}
 
@@
 
+The lemma is proved.
 }
@@
 	At every iteration of the IA-DL method there will exist such interval $[x_{j-1}, x_j]$, $j=j(k)$, that $\bar x \in [x_{j-1}, x_j]$. 
+	{ \color{blue}
 	According to our assumption that $\bar y = y(\bar x)$ is not the limit point, there will exist such number $\bar k$, starting from which the trial points will not fall into the interval $[x_{j-1}, x_j]$.
-That is, to conduct a new trial at the point $x^{k+1} \in [x_{t-1}, x_t]$ for each $k\geq \bar k$ such interval $[x_{t-1}, x_t], t=t(k)$ will be selected that $t(k) \neq j(k)$.
+That is, to conduct a new trial at the point $x^{k+1} \in [x_{t-1}, x_t]$ for each $k\geq \bar k$ such interval $[x_{t-1}, x_t], t=t(k),$ will be selected that $t(k) \neq j(k)$.
+	}
 
@@
 \end{equation}
-{\color{red}
-
+{\color{blue}
 Indeed, both $j$ and $t$ depend from $k$, as the numbering of the trial points (by the lower index) changes at every iteration and the numbering of the intervals changes respectively.
@@
 
-Hence, according to Rules 5 and 6 of the IA-DL method the next iteration falls into the interval $[x_{j-1}, x_j]$, which contradicts our assumption regarding the non-limit nature of the point $\bar x$.
+Hence, according to Steps 5 and 6 of the IA-DL method the next iteration falls into the interval $[x_{j-1}, x_j]$, which contradicts our assumption regarding the non-limit nature of the point $\bar x$.
 The theorem is proven.

### 2020-04-04 2019/NUMTA_Barkalov/SoftComputing/answer.tex
Fix mistakes in English in the answers to reviewer

%% Based on a TeXnicCenter-Template by Gyorgy SZEIDL.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------
%
\documentclass[12pt]{article}%
%
\usepackage{amsmath}%
\usepackage{amsfonts}%
\usepackage{amssymb}%
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{ulem}
\usepackage{cancel}	
%\usepackage[utf8]{inputenc}
%\usepackage[english, russian]{babel} % Русские и английские переносы

%-------------------------------------------

\begin{document}

\section*{Answers to the reviewer's comments}

We are very grateful to the reviewer for his/her useful comments. Please find below our corrections that we suggest to clarify the issues.

\phantom{f}

\textbf{Question 1}. Page 7, Column 1, lines 36--56

... It was shown in (22) the inequality $R_{glob}(i) \geq \Delta_i\left(1-\frac{1}{r_M^{glob}} \right)^2$, 
but in (23) the equality $R_{glob}(i) = \Delta_i \left(1-\frac{1}{r_M} \right)^2$ 
is used both for $r_M = r_{glob}$ and $r_M = r_{loc}$. ... 
This is not sufficient to prove that $R_{glob}(i)\geq R_{loc}(i)$.
In order to obtain (23) and (24) $\mu_M$ has to be equal to 
\[
\mu_M = \frac{\left|z_i-z_{i-1}\right|}{\left(x_i-x_{i-1}\right)^{1/N}}
\]
if $\nu(x_{i-1})=\nu(x_i)=M$. Now can we be sure that this condition is fulfilled?

%\phantom{f}

\textbf{Answer}.
We agree that some of the formulations we used in the paper to prove that $R_{glob}(i)\geq R_{loc}(i)$ were not entirely clear.
We have reformulated the part of the proof on page 7, column 1. 
All corrections have been incorporated in the paper and highlighted in red.

p.7, column 1, lines after formula (21).

If $\nu(x_{i-1}) = \nu(x_i) = M$, then, taking into account (12), the following inequality will be valid
$$
	z_i+z_{i-1}-2z_M^* = |z_i-z_{i-1}| = \mu \Delta_i \leq \mu_M\Delta_i,
$$
where $\mu = |z_i-z_{i-1}|/\Delta_i$.
From here, it will follow that for any value $r_M>1$ the following inequalities will hold
\begin{align}
	R(i) & = \Delta_i + \frac{(z_i-z_{i-1})^2}{r_M^2\mu_M^2\Delta_i} - 2\frac{z_i+z_{i-1}-2z_M^*}{r_M\mu_M} = \nonumber \\
	& = \Delta_i + \frac{\mu^2\Delta_i}{r_M^2\mu_M^2} - 2\frac{\mu\Delta_i}{r_M\mu_M} = \nonumber \\
  & = \Delta_i\left(1 + \left(\frac{\mu}{r_M\mu_M}\right)^2 - 2\frac{\mu}{r_M\mu_M}\right) = \nonumber \\
	& = \Delta_i\left(1 - \frac{\mu}{r_M\mu_M}\right)^2 \geq  \Delta_i \left( 1-\frac{1}{r_M}\right)^2 > 0.\nonumber
\end{align}

And because $r_{M}^{glob} > r_{M}^{loc} > 1$ and $\mu \leq \mu_M$, then
\[ 
	R_{glob}(i) = \Delta_i\left(1-\frac{\mu}{r_M^{glob}\mu_M} \right)^2 > \Delta_i\left(1-\frac{\mu}{r_M^{loc}\mu_M} \right)^2 = R_{loc}(i).
\]

\phantom{f}

\textbf{Question 2}. p. 9 lines 17--22 are not clear.

If $N = 2$ $\max_{0\leq\alpha\leq 1} {\left( \alpha^{\frac{1}{2}}+(1-\alpha)^{\frac{1}{2}} \right)} = \sqrt{2} \neq 2^{3-1/2}=2^{5/2}$ as in line 22. The maximum value is $2^{(N-1)/N}$.

%\phantom{f}

\textbf{Answer}. Yes, the maximum value is $2^{1-1/N}$.
There was an inaccuracy when typing. To correct it , we have removed the extra symbol ``4'' in the formula on p. 9, lines 17--22

\begin{align*}
	...& \geq \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} = \\
	& = \Delta_i\left(1- { \color{red} \cancel 4} \frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0,
\end{align*}

i.e. $4 \cdot 2^{1-1/N} = 2^{3-1/N}$.

\phantom{f}

\textbf{Question 3}. The proof is not self contained, in particular in lines 41--49 the used statements have not appeared before the proof.

%\phantom{f}

\textbf{Answer}. At the beginning of the proof of the theorem we have assumed that $x'$ is a limit point for the IA-DL and  $x' \neq \bar x$. 
This assumption implies that in the method IA-DL in the formula (25) for computing the characteristic, the second component takes the highest values, i.e. $R_{glob}(t) < \rho R_{loc}(t)$. This fact corresponds to using the original index algorithm (IA) with parameters $r_\nu = r_\nu^{loc}, \; 1 \leq \nu \leq m+1$.

In Strongin and Sergeyev (2000) it has been proven that for any limit point $x'$ of the sequence $\{x^k\}$, generated by IA while solving the problem at any $r_\nu > 1, \; 1 \leq \nu \leq m+1$, there exists a sequence
of the intervals $[x_{t-1}, x_t], t=t(k)$, contracting to the point $x'$, i.e. $x' \in [x_{t-1}, x_t], t=t(k)$. Characteristics $R(t)$ of the intervals of the specified sequence will be positive and $R(t) \rightarrow 0$ at $k \rightarrow \infty$. 
Here, we have used this statement without repeating its proof that can be found in Strongin and Sergeyev (2000), p. 523--525.

We have expanded the comment about this on page 9, column 1.

\phantom{f}

\textbf{Question 4}. Both $j$ and $t$ depend from $k$. Although it was proved that $R_{glob}(j) > 0$ for a
fixed $j$, how can we be sure that the condition $R_{glob}(j) > \rho R_{loc}(t)$ is fulfilled for a large enough value of $k$?

%\phantom{f}

\textbf{Answer}.
Yes, both $j$ and $t$ depend from $k$, as the numbering of the trial points (by the lower index) changes at every iteration and the numbering of the intervals changes respectively.

However, number $j(k)$ corresponds to the same interval with fixed characteristic $R_{glob}(j(k))>0$, where $R_{glob}(j(k))= R_{glob}(j(k+1))$, whereas number $t(k)$ corresponds to different intervals with different characteristics $R_{loc}(t(k))$, where $R_{loc}(t(k)) \rightarrow 0$. 
From this we can conclude that the condition $R_{glob}(j) > \rho R_{loc}(t)$ is fulfilled for a large enough value of $k$.

Corresponding comments have been incorporated in the paper on page 9, column 2.

\end{document}

### 2020-04-04 2019/NUMTA_Barkalov/SoftComputing/answer2.tex
Fix mistakes in English in the answers to reviewer

%% Based on a TeXnicCenter-Template by Gyorgy SZEIDL.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------
%
\documentclass[12pt]{article}%
%
\usepackage{amsmath}%
\usepackage{amsfonts}%
\usepackage{amssymb}%
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{ulem}
\usepackage{cancel}	
%\usepackage[utf8]{inputenc}
%\usepackage[english, russian]{babel} % Русские и английские переносы

%-------------------------------------------

\begin{document}

\section*{Answers to the reviewer's comments}

We are very grateful to the reviewer for his/her useful comments. Please find below our corrections that we suggest to clarify the issues.

\phantom{f}

\textbf{Question 1. Comment.} As shown in the last part of the proof for $\nu(x_{i-1}) = \nu(x_i) = M$
\[
\frac{R_{glob}(i)}{R_{loc}(i)}=\left( \frac{1-\frac{\mu}{\mu_M r_M^{glob}}}{1-\frac{\mu}{\mu_M r_M^{loc}}} \right)^2
\]
For these reasons an update of the formula (24) is required.

\textbf{Answer}. Yes, we agree, that 
\[
\frac{R_{glob}(i)}{R_{loc}(i)}=\left( \frac{1-\frac{\mu}{\mu_M r_M^{glob}}}{1-\frac{\mu}{\mu_M r_M^{loc}}} \right)^2 = \left( \frac{1-\frac{1}{r_M^{glob}}}{1-\frac{1}{r_M^{loc}}} \right)^2 = \rho
\]
only if $\mu = \mu_M$.

Accordingly, we have corrected both the formula (24) and the comment following it (see page 7, column 1).
In this comment, we explain that the value 
\[
\rho = \left( \frac{1-\frac{1}{r_M^{glob}}}{1-\frac{1}{r_M^{loc}}} \right)^2
\] 
can be considered as the maximum possible value of the equalization coefficient for $R_{loc}(i)$ and $R_{glob}(i)$.
The use of the equalization coefficient equal to $\rho$ will correspond to equalizing the largest possible difference between the values of $R_{loc}(i)$ and $R_{glob}(i)$ and will contribute to a more local nature of the search when the difference between $R_{loc}(i)$ and $R_{glob}(i)$ is not at its maximum, i.e. when $\mu < \mu_M$.

%To remove all questions мы провели вычислительные эксперименты с использованием в алгоритме выравнивающего  коэффициента $\rho '$ , и они показали худшие результаты, чем использование $\rho$ из .

\phantom{f}

\phantom{f}

\textbf{Question 2. Comment.} Ok

\break

\textbf{Question 3. Comment.} I would suggest to state these results in a preliminary lemma in order
to insert the reference about these in the proof of Theorem 2 replacing the
comments in the lines -11/-1 on p.9 column 1 of the updated paper.

\textbf{Answer}. A corresponding lemma has been added on page 8, column 2, and the reference to it has been inserted in the proof of Theorem 2.

\phantom{f}

\textbf{Question 4. Comment.} I would suggest to explain better the equality $R_{glob}(j(k)) = R_{glob}(j(k+1))$
recalling the assumption that $x'	$ is a limit point for the IA-DL and $x' \neq \bar x$.

\textbf{Answer}. We inserted the suggested notation $\bar k$ on page 9, column 1, for the number of trials starting from which the trial points will not fall into the interval $[x_{j-1}, x_j]$. We also used this notation on page 10, column 1, to explain the inequality $R_{glob}(j) > \rho R_{loc}(t)$.

%\break
%
	%This statement can be interpreted as the incomparability of the local and global characteristics: the global characteristic of the best interval is larger if $\nu(x_{i-1})=\nu(x_i)=M$. At the same time, it follows from the formulae (\ref{R_glob_greater_R_loc}) that 
	%
%\begin{equation}\label{R_glob_equal_R_loc}
	%R_{glob}(i) = \left( \frac{1-\frac{\mu}{\mu_M r_M^{glob}}}{1-\frac{\mu}{\mu_M r_M^{loc}}} \right)^2 R_{loc}(i),
%\end{equation}
%wherein the maximum value of the equalization coefficient is 
%\begin{equation}\label{new_rho}
%\rho=\left(\frac{1-1/r_{M}^{glob}}{1-1/r_{M}^{loc}}\right)^2 
%\end{equation}
%in the case when $\mu = \mu_M$. %, where $\mu = |z_i-z_{i-1}|/\Delta_i$ and $\mu_M$ is from (\ref{current_lower_bounds}).
%The use of the equalization coefficient equal to $\rho$ from (\ref{new_rho}) in all cases will correspond to equalizing the largest possible difference between the values of $R_{loc}(i)$ and $R_{glob}(i)$ and will contribute to a more local nature of the search when the difference between $R_{loc}(i)$ and $R_{glob}(i)$ is not the largest, i.e. when $\mu < \mu_M$.
%
%%Использование нормирующего коэффициента равного $\rho$ из (\ref{new_rho})  будет соответствовать выравниванию наибольшей возможной разницы между значениями $R_{loc}(i)$ и $R_{glob}(i)$ и будет способствовать более ``локальному'' характеру поиска в случае когда разница между $R_{loc}(i)$ и $R_{glob}(i)$ не является максимальной (т.е. когда $\mu_M > \mu$). 
%
%Based on the relations (\ref{R_glob_equal_R_loc}) and (\ref{new_rho}), one can formulate an \textit{index algorithm with dual estimate of the Lipschitz constants} (hereafter referred to as IA-DL).
%

\end{document}

### 2020-04-04 2019/NUMTA_Barkalov/SoftComputing/article.tex
Fix mistakes in English in the answers to reviewer

@@
 \end{equation}
-wherein the maximum value of the equalization coefficient is 
+where the maximum value of the equalization coefficient is 
 \begin{equation}\label{new_rho}
@@
 in the case when $\mu = \mu_M$. %, where $\mu = |z_i-z_{i-1}|/\Delta_i$ and $\mu_M$ is from (\ref{current_lower_bounds}).
-The use of the equalization coefficient equal to $\rho$ from (\ref{new_rho}) in all cases will correspond to equalizing the largest possible difference between the values of $R_{loc}(i)$ and $R_{glob}(i)$ and will contribute to a more local nature of the search when the difference between $R_{loc}(i)$ and $R_{glob}(i)$ is not the largest, i.e.
+The use of the equalization coefficient equal to $\rho$ from (\ref{new_rho}) in all cases will correspond to equalizing the largest possible difference between the values of $R_{loc}(i)$ and $R_{glob}(i)$ and will contribute to a more local nature of the search when the difference between $R_{loc}(i)$ and $R_{glob}(i)$ is not at its maximum, i.e.
 when $\mu < \mu_M$.
@@
 
-Based on the relations (\ref{R_glob_equal_R_loc}) and (\ref{new_rho}), one can formulate an \textit{index algorithm with dual estimate of the Lipschitz constants} (hereafter referred to as IA-DL).
+Based on the relations (\ref{R_glob_equal_R_loc}) and (\ref{new_rho}), one can formulate an \textit{index algorithm with a dual estimate of the Lipschitz constants} (hereafter referred to as IA-DL).
 
@@
 \]
-meaning that the interval $[x_{t-1}, x_t]$ containing the limit point $\bar{x}$ at the step $k$ is hit by the trial at the subsequent step $k+1$.
+meaning that the interval $[x_{t-1}, x_t]$ containing the limit point $\bar{x}$ at the step $k$ is hit in the trial at the subsequent step $k+1$.
 
-From Step 3 and Step 7 of the index algorithm follows that
+From Step 3 and Step 7 of the index algorithm, it follows that
 \begin{equation}\label{max_gamma}
@@
 
-As shown in Lemma 1, at each iteration there exists an interval with strictly positive characteristic. Due to the rule (\ref{MaxR}) this leads to the conclusion that
+As shown in Lemma 1, at each iteration there exists an interval with a strictly positive characteristic. Due to the rule (\ref{MaxR}) this leads to the conclusion that
 intervals from (\ref{intervals_t}) should also have positive characteristics because they are hit by subsequent trials; i.e., the conditions (\ref{R_g_zero}) are necessary.
 
-For intervals with indexes $\nu(x_{i-1}) = \nu(x_{i}) = \nu$ from the second condition of Theorem \ref{theorem:1} and (\ref{K_leq_L}) follows that
+For intervals with indexes $\nu(x_{i-1}) = \nu(x_{i}) = \nu$, it follows from (\ref{current_lower_bounds}) that
 \[
-\frac{\left|z_i-z_{i-1}\right|}{\Delta_i}\leq 2L_\nu\sqrt{N+3},
+\frac{\left|z_i-z_{i-1}\right|}{\Delta_i}\leq \mu_\nu,
 \]
@@
 \[
-z_i = g_\nu(y(x_i))\geq z_\nu^*, \; \nu=\nu(x_i)
+z_i = g_\nu(y(x_i))\geq z_\nu^*, \; \nu=\nu(x_i),
 \]
-holds. Then from the rules (\ref{R_1})--(\ref{R_3}) it follows that $R(i) \leq 2\Delta_i$, if $\nu(x_{i-1}) \neq \nu(x_{i})$, and
+hold. Then from the rules (\ref{R_1})--(\ref{R_3}) it follows that 
 \[
-R(i) \leq \Delta_i\left(1+\left(\frac{2L_\nu\sqrt{N+3}}{r_\nu\mu_\nu}\right)^2\right), 
+R(i) \leq 2\Delta_i, \; \text{if } \; \nu(x_{i-1}) \neq \nu(x_{i}),
 \]
-if $\nu(x_{i-1}) = \nu(x_{i}) = \nu$. These inequalities combined with (\ref{lim_delta_t}) lead us to the validity of (\ref{lim_R}).
-
+\[
+R(i) \leq \Delta_i\left(1+\frac{1}{r_\nu^2}\right), \; \text{if } \; \nu(x_{i-1}) = \nu(x_{i}) = \nu.
+\]
+These inequalities combined with (\ref{lim_delta_t}) lead us to the validity of (\ref{lim_R}).
 The lemma is proved.
@@
 {\color{blue}
-Indeed, both $j$ and $t$ depend from $k$, as the numbering of the trial points (by the lower index) changes at every iteration and the numbering of the intervals changes respectively.
+Indeed, both $j$ and $t$ depend on $k$, as the numbering of the trial points (by the lower index) changes at every iteration and the numbering of the intervals changes respectively.
 %new text begin
-However, $R_{glob}(j(k))>0$ is fixed for each $k \geq \bar k$ because $j(k)$ corresponds to the same interval for $k \geq \bar k$, whereas number $t(k)$ corresponds to different intervals with different characteristics $R_{loc}(t(k)) \rightarrow 0$.
+However, $R_{glob}(j(k))>0$ is fixed for each $k \geq \bar k$ because $j(k)$ corresponds to the same interval for $k \geq \bar k$, whereas the number $t(k)$ corresponds to different intervals with different characteristics $R_{loc}(t(k)) \rightarrow 0$.
 From this we can conclude that the condition $R_{glob}(j) > \rho R_{loc}(t)$ is fulfilled for a large enough value of $k$.

### 2020-04-07 2020/RSD_GPU/RSD2020.tex
Initial tex version of the paper

--- 2020-02-10 2020/PAVT_Nested_Full/pavt_paper.tex
@@
 
-%\pdfmapfile{+txfonts.map}
-
-%\usepackage[utf8]{inputenc}
-%\usepackage[english, russian]{babel} % Русские и английские переносы
-
 % to typeset URLs, URIs, and DOIs
@@
 %
-\title{Combining of local and global search in the parallel nested optimization scheme 
-\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
+\title{Adaptive global optimization using graphics accelerators\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
 }
 %
-\titlerunning{Combining of local and global search}  % abbreviated title (for running head)
+\titlerunning{Adaptive global optimization}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 %
-\author{Konstantin Barkalov\letter\orcidID{0000-0001-5273-2471} \and Ilya 
-Lebedev\orcidID{0000-0002-8736-0652} \and Maria Kocheganova \orcidID{0000-0002-4722-6299} 
-\and Victor Gergel \orcidID{0000-0002-4013-2329}}
+\author{Konstantin Barkalov\letter\orcidID{0000-0001-5273-2471} \and \\  Ilya Lebedev\orcidID{0000-0002-8736-0652}}
 %
-\authorrunning{Konstantin Barkalov et al.} % abbreviated author list (for running head)
+\authorrunning{Konstantin Barkalov \and Ilya Lebedev} % abbreviated author list (for running head)
 %
 %%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{Konstantin Barkalov and Ilya Lebedev, and Maria Kocheganova}
+\tocauthor{Konstantin Barkalov and Ilya Lebedev}
 %
 \institute{Lobachevsky State University of Nizhni Novgorod, Russia  \\
-	\email{konstantin.barkalov@itmm.unn.ru}
+	\email{konstantin.barkalov@itmm.unn.ru},
+	\email{ilya.lebedev@itmm.unn.ru}
 }
@@
 
-The present paper considers global optimization problems and the numerical methods of solving them.
-An assumption is made regarding the dependence of the objective function on its parameters: it is multiextremal with respect to selected variables only, and the dependence on the other variables has a local character.
-Problems of this type may arise in identifying the unknown parameters of the mathematical models, based on experimental results.
-A parallel computational scheme, which accounts for this feature, is being proposed.
-This novel scheme is based on the idea of nested (recursive) optimization, when at the upper recursion level the optimization with respect to the significant variables is performed (using the global optimization method), and at the lower one local optimization problems are solved (using the local method).
-The generated local subproblems will not affect each other, and solving these problems can be performed in parallel.
-Numerical experiments have been conducted on several hundred test problems, confirming the efficiency of the proposed scheme of parallel computations.
-
-\keywords{Global optimization $\cdot$ Local optimization $\cdot$ Recursive optimization scheme 
-$\cdot$ Parallel algorithms}
+\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Reduction of dimensionality $\cdot$ Peano space-filling curves $\cdot$ Recursive optimization $\cdot$ Graphics accelerators }
 \end{abstract}
@@
 
-Global optimization problems arise frequently in various fields of modern science and technology.
-For example, problems of finding the configurations of various chemical compounds corresponding to the minimum interaction energy are relevant \cite{Posypkin2014}.
-In particular,  such problems arise
-in the development of new medical drugs \cite{Sulimov}. 
-Another example is the problem of finding optimal control parameters for complex processes \cite{Kalyulin2017,Modorskii2016}.
-Global optimization problems arise in many other applications as well (see the review \cite{Pinter2006}).
-
-Application of global optimization methods for identifying the parameters of the mathematical models based on the data of experiments has become a classical practice.
-Problems of this type require a search for the values of the unknown parameters of the model, from which computational results according to the model are close to the results obtained experimentally.
-
-The number of parameters which are required to be identified in such a way, may reach tens and hundreds in complex mathematical models \cite{Akhmadullina2017,Nurislamova2016}.
-In problems of such dimensionality, regular search methods for global solution cannot be applied because of the extremely large computational costs for covering the search domain by the trial points.
-This remains true even with the use of efficient algorithms (for example, \cite{Evtushenko2009,Jones2009,Paulavicius2011}) constructing essentially nonuniform coverages.
-
-However, in the problems of model identification, the parameters, which will affect the result most significantly, can often be selected.
-As a rule, the number of such parameters is small.
-The rest of the parameters either cause an insufficient effect (in this case, multiextremality can be neglected) or dependence on the second group of parameters has a local character.
-
-So far, the development of parallel algorithms for minimizing the essentially multidimensional functions (hundreds of variables), which accounts for the dependence of the objective function on different groups of parameters, is relevant.
-Specifically, the objective function is assumed to be multiextremal with respect to some of the variables.
-The rest of the variables have a local effect, i.e., the objective function is unimodal with respect to them.
-In this case, solving the problem can be organized according to the nested (recursive) optimization scheme.
-Solving the multiextremal subproblem, which requires the use of complex global search algorithms, will be performed at the upper level of recursion.
-The unimodal subproblems (each corresponding to a fixed set of values of the first part of parameters) will be solved at the lower level.
-For solving these subproblems, efficient methods of local optimization, linear programming, or linear algebra can be applied (subject to the character of the local subproblems).
-
-The text describing the results of the present study is organized as follows.
-Section 2 describes the nested optimization scheme taking into account different properties of the objective function at different nesting levels.
-A particular example of a problem of this class is presented, along with a general scheme of the organization of parallel computations.
-Section 3 is devoted to the description of the parallel global search algorithm using space-filling curves.
-The computational scheme of the parallel algorithm is presented, and its main properties are described.
-Section 4 discusses the methods of solving the local subproblems within the framework of the nested optimization scheme.
-A brief description of the properties of the local extremum search problems and known methods of solving them is given.
-Section 5 contains the results of a numerical experiment.
-The evaluation of the scalability of the proposed parallel computation scheme in solving a series of test problems was performed.
-Section 6 concludes the paper.
-
 \section{Nested optimization scheme}
-
-Let us consider an optimization problem of the type
-\begin{eqnarray}\label{main_problem}
-& f(x^\ast,y^\ast)=\min{\left\{f(x,y):x\in S, y\in D\right\}}, \nonumber \\
-& S=\left\{x\in R^M: a_i\leq x_i \leq b_i, 1\leq i \leq M\right\}, \\
-& D=\left\{y\in R^N: c_i\leq y_i \leq d_i, 1\leq i \leq N\right\}. \nonumber
-\end{eqnarray}
-
-Let us assume that at any fixed set of the values $\overline{x}$ the function $f(\overline{x},y)$ is a 
-multiextremal one and satisfies Lipschitz condition
-\[
-\left|f(\overline{x},y')-f(\overline{x},y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in D,\; 0<L<\infty,
-\]
-with the constant $L$ unknown a priori.
-At the same time, at any fixed set of values $\overline{y}$ the function $f(x,\overline{y})$ is a 
-unimodal one, i.e. the function $f(x,\overline{y})$ has a single minimum point $x^*$ (depending, in 
-general, on $\overline{y}$).
-
-Accounting for such a peculiarity of the considered problem class can reduce essentially the computational costs 
-of the optimum search. Indeed, according to known nested optimization scheme 
-\cite{Carr}, the solving of the initial problem can be reduced to the problem of finding the global 
-minimum of the function $\varphi(y)$
-\begin{equation}\label{global_problem}
-\varphi(y^*) = \min_{y\in D}  \varphi (y),
-\end{equation}
-where 
-\begin{equation}\label{local_problem}
-\varphi(y) = \min_{x\in S} f(x,y).
-\end{equation}
-According to (\ref{global_problem}), the computing of a single value of the function $\varphi (y)$ 
-(let us call this process a search trial) implies the solving of a unimodal problem (\ref{local_problem}), 
-which can be performed by one of the local search methods. The efficiency and sophistication of the 
-local optimization methods allow solving the subproblems (\ref{local_problem}) with the precision 
-exceeding the one of the global optimization methods essentially. Correspondingly, the problem 
-(\ref{global_problem}) can be considered as a global optimization problem, in which the objective 
-function value can be computed with high precision but this operation is time-consuming.
-
-The following approximation problem can be considered as a particular example of a problem, in 
-which different character of dependence on different groups of parameters takes place. 
-Let us assume $m$ values $u_j = u(t_j), 1 \leq j \leq m, $ of the function $u(t)$ to have been obtained 
-in the experiment and the analytical expression for $u(t)$ is known to take the form 
-\begin{equation}\label{ex_func}
-u(t) = u(t,\omega,c)=\sum^{n}_{i=0}\left[c_{2i+1}\sin(\omega_it) + c_{2i+2}\cos(\omega_it)\right].	
-\end{equation}
-In this expression, $c=\left(c_1,\ldots,c_{2n+2}\right)$, 
-$\omega=\left(\omega_0,\omega_1,\ldots,\omega_n\right)$ are the unknown parameters defining 
-particular function $u(t)$.
-
-Let us introduce a measure of the deviation of the function $u(t,\omega,c)$ from the experiential data as a 
-sum of squares 
-\[
-\Delta(\omega,c)= \sum^{m}_{j=1}\left[u_j-u(t_j,\omega,c)\right]^2. 
-\]
-Now, following the idea of least squares fitting, it is possible to present the approximation problem as 
-the problem of minimizing the objective function
-\begin{equation}\label{ex_prob}
-\Delta(\omega^*,c^*) = \min_{\omega,c} \Delta(\omega,c).	
-\end{equation}
-The solution $(\omega^*,c^*)$ of this problem will define the best approximation.
-
-The problem (\ref{ex_prob}) can be written in a recursive form 
-\begin{equation}\label{outer_prob}
-\varphi(\omega^*) = \min_\omega \varphi(\omega),
-\end{equation}
-\begin{equation}\label{inner_prob}
-\varphi(\omega) = \min_c \Delta(\omega,c).
-\end{equation}
-The nested subproblem (\ref{inner_prob}) is a classical linear least squares problem, and its solution 
-can be obtained by solving a system of linear algebraic equations regarding the unknown $c$, which 
-can be done e.g. by Gaussian elimination. At the same time, the external problem (\ref{outer_prob}) 
-will be, in general, a multiextremal one, and its solving will require considerable 
-computational resources.
-
-So far, to solve the problem (\ref{main_problem}) in the nested scheme (\ref{global_problem}), 
-(\ref{local_problem}) one can apply parallel global search algorithm for solving the external 
-subproblems (\ref{global_problem}). At that, a set of independent subproblems (\ref{local_problem}) 
-will be generated at every iteration of the algorithm, the solving of which can be performed in parallel 
-using the local methods.
-A general scheme of the organization of the computations using several processors is presented in 
-Fig.~\ref{MPI_tree}.
-
-\begin{figure}
-\center{\includegraphics[width=1.0\linewidth]{MPI_Tree.pdf}}
-\caption{A tree of subproblems in the parallel nested optimization scheme}
-\label{MPI_tree}
-\end{figure}
-
-The processes of the parallel program will form a tree. 
-The root process will solve the problem (\ref{global_problem}) and distribute the subproblems 
-(\ref{local_problem}) among the child processes. Each subproblem is solved by a separate process, 
-the data exchange between the processes can be organized using MPI. The data transmission between 
-the processes will be minimal: it is required to send the coordinates of points $y^1, ..., y^p$ from the 
-root to the child processes and to receive back the found minimal values 
-\[
-\varphi(y^1) = \min_x f(x,y^1),\; ..., \; \varphi(y^p) = \min_x f(x,y^p).
-\]
-
-A detailed discussion of the algorithms used at different levels of the process tree is given in the next 
-sections.
 
@@
 
-An efficient parallel global optimization algorithm intended for solving the problems of the kind 
-(\ref{global_problem}) has been developed in Lobachevsky University of Nizhni Novgorod (UNN) 
-\cite{Barkalov2018,globalizerSystem,Strongin2018}. 
-The main idea of the parallelization consists in such of organization of computations that several trials 
-are performed in parallel. This approach is featured by high efficiency (the part of the computational 
-process is parallelized, in which the major amount of computations is performed) and generality (it is 
-applicable for a wide class of algorithms).
-
-In the developed global search algorithm, a novel method for the reduction of the dimensionality of 
-the problems to be solved is applied. 
-The reduction of the dimensionality is based on a fundamental fact, according to which an 
-$N$-dimensional hyperparallelepiped $D$ and the interval $[0,1]$ of the real axis are the 
-equal-power sets, and the interval $[0,1]$ can be mapped continuously into $D$ using the Peano 
-curve $y(x)$, i.e. $D = \left\{y(x):0\leq x\leq 1\right\}$.
-
-Utilizing this fact, one can reduce the solving of the problem of minimization of a multidimensional 
-function $\varphi(y)$ to the minimization of a one-dimensional function $\varphi(y(x))$
-\begin{equation}\label{1d_problem}
-\varphi(y (x^*)) = \min \left\{ \varphi (y(x)) : x\in [0,1] \right\},
-\end{equation}
-where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
-\[
-\left|\varphi(y(x'))-\varphi(y(x''))\right|\leq H\left|x'-x''\right|^{1/N}
-\]
-with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation
-$ H=2 L \sqrt{N+3}$ and $y(x)$ is the Peano curve from $[0,1]$ onto $D$.
-
-The algorithms for the numerical construction of the Peano curve approximations (called 
-\textit{evolvents})
-are given in \cite{Sergeyev2013,Strongin2000}. As an illustration, two evolvents are shown in 
-Fig.~\ref{fig_evolvent}. Figure demonstrates that the precision of the evolvent is depend on the 
-density level $m$ used in the construction.
-
-\begin{figure}
-\center
-\begin{minipage}{0.48\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (a)}
-\end{minipage}
-\begin{minipage}{0.48\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (b)}
-\end{minipage}
-\caption{Evolvents in two dimensions with (a) $m=4$ and (b) $m=5$}
-\label{fig_evolvent}
-\end{figure}
-
-It is worth noting that many well known global optimization algorithms are based implicitly on the 
-idea of the dimensionality reduction and adaptation of the one-dimensional algorithms for solving the 
-multidimensional problems \cite{Evtushenko2013,Sergeyev2006,Zilinskas2008}.
-
-According to \cite{Strongin2018,Strongin2000}, the rules of the global search algorithm, in which 
-$p$ trials are performed in parallel at every iteration, are as follows.
-At the first iteration of the method, the trials are performed in parallel in two boundary points $x^1 = 
-0, \; x^2 = 1$ as well as in $(p-2)$ arbitrary internal points of the interval $[0,1]$, i.e. 
-$x^i\in(0,1),3\leq i \leq p$.
-
-Suppose $n\geq 1$  iterations of the method have been carried out, in which trials were performed at 
-$k=k(n)$ points $x^i,1\leq i \leq k$. Then points $x^{k+1},\dots,x^{k+p}$  of search trials of the next 
-iteration are determined according to the following rules.
-
-Rule 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the
-lower indices in ascending order of coordinate values, i.e.
-\[
-0=x_1<x_2<\dots <x_{k-1} <x_k=1,
-\]
-and juxtapose to them the values $z_i=\varphi(y(x_i)), \; 1 \leq i \leq k$, computed at these points.
-
-Rule 2. Compute the current lower estimates
-\begin{equation}\label{Rule_Mu}
-\mu = \max\left\{ \frac{\left|z_i-z_{i-1}\right|}{ \Delta_i },\; 2 \leq i \leq k  \right\},
-\end{equation}
-where $\Delta_i = (x_i-x_{i-1})^{1/N}$. If $\mu$ is equal to zero, then assume $\mu = 1$.
-
-Rule 3. For each interval ($x_{i-1},x_i), \; 2 \leq i \leq k,$ compute
-the \textit{characteristics} $R(i)$ :
-\begin{equation}\label{Rule_R}
-R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r^2 \mu^2\Delta_i}-2\frac{z_i+z_{i-1}}{r \mu},
-\end{equation}
-where $\Delta_i=(x_i-x_{i-1})^{1/N}$. The value $r > 1$ is parameter of the algorithm. An 
-appropriate selection
-of $r$ allows to consider the product $r \mu$ as an estimate
-of the Lipschitz constant $L$ of the objective function.
-
-Rule 4. Arrange characteristics  $R(i), 2 \leq i \leq k$, in decreasing order 
-\begin{equation}\label{Rule_Max}
-R(t_1)\geq R(t_2)\geq \dots \geq R(t_{k}) \geq R(t_{k})
-\end{equation}
-and select $p$ maximum characteristics with interval numbers $t_j, 1\leq j \leq p$.
-
-Rule 5. Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), \; 1\leq j\leq p$, calculated 
-using the formulae
-\begin{equation}\label{Rule_X}
-x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - 
-\frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^
-N.
-\end{equation}
-
-The algorithm terminates if the condition $\Delta_{t_j}<\epsilon$ is satisfied at least for one number 
-$t_j, 1 \leq j \leq p$ ; $\epsilon>0$ is the predefined accuracy.
-
-This method of parallelization has the following justification. The characteristics of intervals 
-(\ref{Rule_R}) used in the method can be considered as probability measures of the global minimum 
-point location in these intervals. Inequalities (\ref{Rule_Max}) arrange intervals according to their 
-characteristics, and trials are carried out in parallel in the first $p$ intervals with the largest 
-probabilities.
-
-Various modifications of this algorithm and the corresponding theory of
-convergence are given in \cite{Barkalov2018,Sergeyev2013,Strongin2018,Strongin2000}.
-%Замечание по поводу теоретических свойств и сравнения алгоритмов
-The results of the experimental comparison  (see, for example, \cite{Sovrasov2019}) show that global search algorithm is superior to many known methods of similar purpose, including both deterministic and heuristic  algorithms.
-
-\section{Local search algorithms}
-
-In the present section, let us consider the issues related to the solving the subproblem 
-(\ref{local_problem}), which (in the notation $f(x) = f(x,\overline{y})$ for the fixed value 
-$\overline{y} \in D$) corresponds the problem of finding a local extremum
-
-\begin{equation} \label{lp}
-f(x) \rightarrow \min, x\in S. 
-\end{equation}
-
-To date, a huge number of various local optimization methods for the problems of the kind (\ref{lp}) 
-have been developed. Among these ones, there are, for example, the gradient methods, Newton and 
-quasi-Newton methods, the methods of conjugate directions, etc. The majority of these ones utilize 
-the principle of local descent, when the method passes to the points with smaller values of the 
-objective function at every step progressively. Almost all these methods can be represented in the 
-form of the iterative relation 
-\[
-x^{k+1} = x^k + s^k d^k,
-\]
-where $x^k$ are the points of main trials consisting in computing a set of some local characteristics of 
-the objective function $I^k=I(x^k)$ at the point $x^k$, where $d^k$ are the directions of shift from 
-the points $x^k$ computed on the base of the results of main trials and $s^k$ are the coefficients 
-defining the magnitudes of the shifts along to the directions chosen.
-
-In order to define the shift magnitudes $s^k$ along the directions $d^k$, the methods can perform 
-auxiliary (working) steps. This results in additional measuring of the local characteristics of the 
-objective function along the direction $d^k$. The transitions from the points $x^k$ to the points 
-$x^{k+1}$ is performed in such a way that an essential decrease of the value of the function $f^k = 
-f( x^k )$ is provided as a result of the step.
-
-The set of the computed local characteristics $I^k=I(x^k)$ may include: the function values $f^k = 
-f( x^k )$, the gradient $\nabla f^k = \nabla f(x^k)$, the matrix of the second derivatives 
-(Hessian) $\Gamma_k=\Gamma^f(x^k)$. Particular set of the characteristics to be measured depends 
-on the properties of the problem being solved as well as on the optimization method chosen.
-
-In the local optimization problems arising in the applications, the a priori information on the objective 
-functions is usually quite limited. For example, a local character of the dependence of the function on 
-the parameters may be suggested only whereas the gradient values (and, moreover, Hessian) are
-unknown. In this case, the direct search methods are applied, which don't utilize any suggestions on
-the smoothness of the objective function. When searching for the minimum, these methods measure 
-the function values only. The rules of the placement of the iteration points in these ones is based on 
-some heuristic logical schemes.
-
-Hooke--Jeeves method \cite{HookJeeves} and Nelder--Mead method \cite{NelderMead} belong to 
-the popular methods of the direct search. In spite of apparent simplicity and theoretical 
-non-substantiation of the direct search methods, these ones are well established in the practical 
-computations. This can be explained as follows. Many methods of smooth optimization are very 
-sensitive to the computational errors in the function values transforming theoretically smooth function 
-into the non-smooth one actually. Because of this, in the practical computations these ones loose the 
-positive properties, which are promised by theory often. The use of the direct search methods allows 
-achieving better results in these conditions.
-
 \section{Numerical experiments}
-
-The scheme of parallel computations described in previous sections was implemented in the 
-Globalizer software system developed in UNN \cite{globalizerSystem,Sysoyev2017}.  
-The numerical experiments were carried out using Lomonosov supercomputer (Lomonosov Moscow 
-State University). Each supercomputer node included two quad-core processors Intel Xeon X5570 
-2.93GHz and 12 Gb RAM. When building the Globalizer system for running on the Lomonosov 
-supercomputer GCC 5.5.0 compiler and Intel MPI 2017 were used.
-
-In order to simulate the applied problems, in which the locally-affecting parameters can be selected, 
-the test function of the kind 
-\begin{equation}\label{test_problem}
-f(x,y) = G(y)+R(x)
-\end{equation}
-were used when conducting the experiments. Here $G(y)$ is a multiextremal function of the 
-dimensionality $N=2$ and $R(x)$ is a unimodal function of the dimensionality $N \gg 2$.
-As the multiextremal part of the problem, the functions 
-\begin{eqnarray} \nonumber \label{vagris}
-G(y)= 
--&\left\{\left(\sum^{7}_{i=1}\sum^{7}_{j=1}A_{ij}g_{ij}(y)+B_{ij}h_{ij}(y)\right)^2+\right. \\
-&\left.\left(\sum^{7}_{i=1}\sum^{7}_{j=1}C_{ij}g_{ij}(y)+D_{ij}h_{ij}(y)\right)^2\right\}^{1/2
-},\\ \nonumber
-\end{eqnarray}
-were considered, where
-\begin{eqnarray} \nonumber
-& y=(y_1,y_2)\in R^2, 0 \leq y_1,y_2 \leq 1, \\ \nonumber
-& g_{ij}(y)=\sin(i\pi y_1)\sin(j\pi y_2),  \\ \nonumber
-& h_{ij}(y)=\cos(i\pi y_1)\cos(j\pi y_2), \nonumber 
-\end{eqnarray}
-and the coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$ are the random numbers distributed uniformly 
-in the interval $[-1,1]$.
-This class of functions is often used for testing the global optimization algorithms. 
-As the local part of the problem, the modified Rosenbrock function 
-\[
-R(x)= \sum_{i=1}^{N-1}{\left[(1-x_i)^2+100(x_{i+1}-x_i^2)^2\right]}, -2 \leq x_i \leq 2 , 1\leq 
-i\leq N,
-\]
-was used, in which the extremum point was shifted in the search domain randomly using the linear 
-coordinate transform.
-
-As an example, a contour plot of a pair of functions  $G(y)$ and $R(x)$ is presented in 
-Fig.~\ref{fig_level}(a),(b). These functions are the complex ones for corresponding global and local 
-optimization methods since the function $G(y)$ is essentially multiextremal one, and the function 
-$R(x)$ has a ravine structure manifested clearly.
-
-\begin{figure}
-\center
-\begin{minipage}{0.48\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{grishagin.png} \\ (a)}
-\end{minipage}
-\begin{minipage}{0.48\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{rosenbrock.png} \\ (b)}
-\end{minipage}
-\caption{Contour plot of the functions (a) $G(y)$ and (b) $R(x)$}
-\label{fig_level}
-\end{figure}
-
-In Fig.~\ref{fig_level}, the trial points performed by the global search algorithm and Hooke--Jeeves 
-method when solving corresponding problems are marked also. 
-The positions of the trial points demonstrate clearly the multiextremality of the function $G(y)$ (the 
-trial points form a nonuniform coverage of the search domain) as well as the ravine character 
-of the function $R(x)$ (the trial points form an elongated group along the ravine bottom).
-
-The problem was considered to be solved if in the course of minimization of the function $G(y)$, the 
-global search algorithm generates the next trial point $y^k$ in the $\epsilon$-nearness of the global 
-minimum $y^*$, i.e. $\left\|y^* - y^k\right\|\leq\epsilon$, where $\epsilon = 10^{-2}$. 
-At that, the precision of finding the minimum when minimizing the unimodal function $R(x)$ was 
-selected 1000 times lower, i.e. $\delta = 10^{-5}$.
-In the global search algorithm, the parameter $r=4.5$ from (\ref{Rule_R}) and the evolvent building 
-parameter $m=10$ were used; the local Hooke--Jeeves method employed has no additional parameters.
-
-In the first series of experiments, 100 problems of the kind (\ref{test_problem}) were solved, in which 
-the dimensionality of the local subproblems was $N=50$. The problems were solved using 1, 4, and 8 
-cluster nodes, in which from 2 up to 32 MPI-processes were employed. According to the 
-parallelization scheme presented in Fig.~\ref{MPI_tree}, the solving of the global optimization 
-problem at the upper nesting level was performed in a single (master) process, which initiated the 
-parallel solving of the local problems of the lower level (from 1 up to 31 problems, respectively). So 
-far, the minimum possible number of processes employed was equals to 2, one per each nesting level.
-
-Table \ref{tab1} presents the averaged number of iterations $K_{av}$ performed by the global search 
-algorithm when minimizing the function $G(y)$, the averaged time of solving the whole problem 
-$T_{av}$, and the speed-up in time $S$ subject to the number of employed MPI-processes $p$.
-
-\begin{table}
-\centering
-\caption{Results of solving a series of problems of dimensionality $N=52$}\label{tab1}
-\begin{tabular}{cccc}
-\hline\noalign{\smallskip}
- $\;\;\;p\;\;\;$  &  $\;\;\;K_{av}\;\;\;$ &  $\;\;\;T_{av}\;\;\;$ & $\;\;\;S\;\;\;$ \\
-\noalign{\smallskip}\hline\noalign{\smallskip}
- 2  & 349.4  & 9.14 & ---  \\
- 16 & 25.6   & 0.70 & 13.2 \\
- 32 & 12.9   & 0.37 & 24.6 \\
-\noalign{\smallskip}\hline
-\end{tabular}
-\end{table}
-
-In the second series of experiments, 10 problems of the kind (\ref{test_problem}) were solved, in 
-which the dimensionality of the local subproblems was equal $N=100$. The running parameters were 
-the same as in the previous series of experiments. The results of solving are presented in Table 
-\ref{tab2}.
-
-\begin{table}
-\centering
-\caption{Results of solving a series of problems of dimensionality $N=102$}\label{tab2}
-\begin{tabular}{cccc}
-\hline\noalign{\smallskip}
- $\;\;\;p\;\;\;$  &  $\;\;\;K_{av}\;\;\;$ &  $\;\;\;T_{av}\;\;\;$ & $\;\;\;S\;\;\;$ \\
-\noalign{\smallskip}\hline\noalign{\smallskip}
- 2  & 299.1  & 41.19 & ---  \\
- 16 & 22.0   & 2.92 & 14.1 \\
- 32 & 11.7   & 1.50 & 27.4 \\
-\noalign{\smallskip}\hline
-\end{tabular}
-\end{table}
-
-% В последней серии экспериментов было решено 10 задач, в которых в качестве многоэкстремальной части использовались трехмерные задачи из класса GKLS.
-Данный класс задач часто используется для тестирования методов глобальной оптимизации [ссылки].
-Размерность локальных подзадач была равна N = 100.
-Задачи решались на 1, 8, 16 и 32 узлах кластера, на которых использовалось от 4 до 128 MPI-процессов.
-Параметры запуска такие же, как в предыдущих сериях экспериментов.
-Результаты решения представлены в таблице, ускорение посчитано относительно запуска на одном узле на 4 mpi процесса.
-
-In the last series of experiments, 10 problems were solved, in which three-dimensional functions from GKLS class \cite{Gaviano2003} were used as a multiextremal part. 
-This class of functions is often used for investigating global optimization algorithms \cite{Barkalov2018,Paulavicius2014,Sergeyev2015}.
-The dimensionality of the local subproblems was equal $N=100$.
-The problems were solved at 1, 8, 16 and 32 cluster nodes, on which from 4 to 128 MPI processes were used.
-The running parameters were the same as in the previous series of experiments.
-The results of solving are presented in Table \ref{tab3}, where speedup is calculated with respect to  running on a single node using 4 MPI processes.
-
-%In the last series of experiments, 10 problems of the kind (\ref{test_problem}) were solved, in which the dimensionality of the local subproblems was equal $N=100$.
-The running parameters were the same as in the previous series of experiments.
-The results of solving are presented in Table \ref{tab3}.
-
-\begin{table}
-\centering
-\caption{Results of solving a series of problems of dimensionality $N=103$}\label{tab3}
-\begin{tabular}{cccc}
-\hline\noalign{\smallskip}
- $\;\;\;p\;\;\;$  &  $\;\;\;K_{av}\;\;\;$ &  $\;\;\;T_{av}\;\;\;$ & $\;\;\;S\;\;\;$ \\
-\noalign{\smallskip}\hline\noalign{\smallskip}
- 4  & 730.8  & 63.4 & ---  \\
- 32 & 87.4   & 7.37 & 8.6 \\
- 64 & 37.0   & 3.77 & 16.8 \\
- 128 & 16.5   & 1.78 & 35.7 \\
-\noalign{\smallskip}\hline
-\end{tabular}
-\end{table}
 
@@
 
-Results of experiments performed on the series of optimization problems demonstrate that the scheme of parallel computations proposed in the present work has good potential for parallelization.
-Specifically, when solving a series of 100 problems with a total number of variables equal to 52 (from which the two first variables caused a global effect on the objective function and fifty affected the one locally), a speedup of 24.6 was obtained when using 32 processes.
-When solving  problems with more complex local subproblems (with the number of the local variables N=100), the speedup increased up to 35 when using 128 processes.
-In the last case the speedup was calculated with respect to the use of 4 processes on one node.
-The developed scheme of parallel computations can be applied in identifying complex mathematical models, which are featured by large numbers (tens and hundreds) of unknown parameters.
+Simple test of new references
+\cite{Hwu2011,Ferreiro2013,Garcia2014,Langdon2011,He2008,Evtushenko2009,Paulavicius2011,Kindratenko2014}
+Almost all the old references are already in the bib-file.
+
 %

### 2020-04-08 2019/NUMTA_Barkalov/SoftComputing/article.tex
Fix final reviewr's comments

@@
 The upper estimate provides the global convergence whereas the lower one reduces the number of trials necessary to find the global optimizer with the required accuracy.
-The considered algorithm for solving the constrained problems doesn’t use the ideas of the penalty function method; each constraint of the problem is accounted for separately.
+The considered algorithm for solving the constrained problems doesn't use the ideas of the penalty function method; each constraint of the problem is accounted for separately.
 The convergence conditions of the proposed algorithm are formulated in the corresponding theorem.
@@
 where $\gamma = \frac{r+1}{2r}<1$  and $ r=\min\{r_\nu:1\leq\nu\leq m+1\}>1$.
-The contraction property (\ref{lim_delta_t}) and the property (\ref{nested_int}) for the product of all intervals (\ref{intervals_t}) are the consequences of (\ref{max_gamma}).
+The contraction property (\ref{lim_delta_t}) and the property (\ref{nested_int}) for the intersection of all intervals (\ref{intervals_t}) are the consequences of (\ref{max_gamma}).
 

### 2020-04-12 2020/RSD_GPU/RSD2020.tex
Introduction

@@
 
+Problems of multidimensional multiextremal optimization and numerical methods for their solution are considered.
+The general assumption is made about the function being optimized: it satisfies the Lipschitz condition with an a priori unknown constant.
+Many approaches to solving problems of this class are based on reducing the dimension of the problem; i.e., addressing a multidimensional problem by solving a family of problems with lower dimension.
+In this work, an adaptive dimensionality reduction scheme is investigated, and its implementation using graphic accelerators is proposed.
+Numerical experiments on several hundred test problems were carried out, and they confirmed acceleration in the developed GPU version of the algorithm.
+
 \keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Reduction of dimensionality $\cdot$ Peano space-filling curves $\cdot$ Recursive optimization $\cdot$ Graphics accelerators }
@@
 
-\section{Nested optimization scheme}
+A promising direction in the field of parallel global optimization (which, indeed, is true in many areas related to the software implementation of time-consuming algorithms) is the use of graphics processing units (GPUs).
+In the past decade, graphics accelerators have rapidly increased performance to meet the ever-growing demands of graphics application developers.
+Additionally, in the past few years some principles for developing graphics hardware have changed, and as a result it has become more programmable.
+Today, a graphics accelerator is a flexibly programmable, massive parallel processor with high performance, which is in demand for solving a range of computationally time-consuming problems \cite{Hwu2011}.
+
+However, the potential for graphics accelerators to solve global optimization problems has not yet been fully realized.
+Using the GPUs, they basically parallelize nature-inspired optimization algorithms, which are somehow based on the idea of random search (see, for example, \cite{Ferreiro2013,Garcia2014,Langdon2011}).
+By virtue of their stochastic nature, algorithms of this type guarantee convergence to the global minimum only in the sense of probability, which differentiates them unfavorably from deterministic methods.
+
+With regard to many deterministic algorithms of Lipschitzian global optimization with guaranteed convergence, parallel variants have been proposed \cite{Evtushenko2009,He2008,Paulavicius2011}.
+However, these versions of algorithms are parallelized on the CPU using shared and/or distributed memory; presently, no GPU implementations have been made.
+For example, \cite{Paulavicius2011} describes parallelization of an algorithm based on the ideas of the branch and boundary method using MPI and OpenMP.
+
+Within the framework of this research, we consider the problems of capturing the optimum, which are characterized by a lengthy period for calculating the values of objective function in comparison with the time needed for processing them.
+For example, objective function can be specified using systems of linear algebraic equations, systems of ordinary differential equations, etc.
+Currently, graphics accelerators can be used to solve problems of this type.
+Moreover, an accelerator can solve several such problems at once \cite{Kindratenko2014}; i.e., using the GPU, one can calculate multiple function values simultaneously.
+
+Thus, calculating the optimization criterion can be implemented on the GPU, and the role of the optimization algorithm (running on the CPU) consists in the effective selection of points for conducting parallel tests.
+This scheme of working with the accelerator is fully consistent with the work of the parallel global search algorithm developed at the Lobachevsky State University of Nizhni Novgorod and presented in a series of publications [references].
+
+\section{Multidimensional parallel global search algorithm}
 
 \section{Parallel global search algorithm}
+
+\section{Dimensionality reduction schemes in global optimization problems}
+
+\subsection{Dimensionality reduction using multiple mappings}
+
+\subsection{Adaptive dimensionality reduction scheme}
+
+\section{GPU implementation}
+
+\subsection{General scheme}
+
+\subsection{Organization of parallel computing}
 
@@
 
-Simple test of new references
-\cite{Hwu2011,Ferreiro2013,Garcia2014,Langdon2011,He2008,Evtushenko2009,Paulavicius2011,Kindratenko2014}
-Almost all the old references are already in the bib-file.
+In summary, we observe that the use of graphics processors to solve global optimization problems shows noteworthy promise, because high performance in modern supercomputers is achieved (mainly) through the use of accelerators.
+
+In this paper, we consider a parallel algorithm for solving multidimensional multiextremal optimization problems and its implementation on the GPU.
+In order to experimentally confirm the theoretical properties of the parallel algorithm under consideration, computational experiments were carried out on a series of several hundred test problems of different dimensions.
+The parallel algorithm demonstrates good acceleration, both in the GPU and CPU versions.
 

### 2020-04-12 2020/RSD_GPU/RSD2020.tex
Первая версия, переведенного текста в tex.

@@
 \usepackage{marvosym}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage[russian]{babel}
 
@@
 
-\maketitle              % typeset the title of the contribution
+%\maketitle              % typeset the title of the contribution
 
@@
 Additionally, in the past few years some principles for developing graphics hardware have changed, and as a result it has become more programmable.
-Today, a graphics accelerator is a flexibly programmable, massive parallel processor with high performance, which is in demand for solving a range of computationally time-consuming problems \cite{Hwu2011}.
+Today, a graphics accelerator is a flexibly programmable, massive parallel processor with high performance, which is in demand for solving a range of computationally time-consuming problems \cite{Ref1}.
 
 However, the potential for graphics accelerators to solve global optimization problems has not yet been fully realized.
-Using the GPUs, they basically parallelize nature-inspired optimization algorithms, which are somehow based on the idea of random search (see, for example, \cite{Ferreiro2013,Garcia2014,Langdon2011}).
+Using the GPUs, they basically parallelize nature-inspired optimization algorithms, which are somehow based on the idea of random search (see, for example, \cite{Ref2,Ref3,Ref4}).
 By virtue of their stochastic nature, algorithms of this type guarantee convergence to the global minimum only in the sense of probability, which differentiates them unfavorably from deterministic methods.
 
-With regard to many deterministic algorithms of Lipschitzian global optimization with guaranteed convergence, parallel variants have been proposed \cite{Evtushenko2009,He2008,Paulavicius2011}.
+With regard to many deterministic algorithms of Lipschitzian global optimization with guaranteed convergence, parallel variants have been proposed \cite{Ref5,Ref6,Ref7}.
 However, these versions of algorithms are parallelized on the CPU using shared and/or distributed memory; presently, no GPU implementations have been made.
-For example, \cite{Paulavicius2011} describes parallelization of an algorithm based on the ideas of the branch and boundary method using MPI and OpenMP.
+For example, \cite{Ref7} describes parallelization of an algorithm based on the ideas of the branch and boundary method using MPI and OpenMP.
 
@@
 Currently, graphics accelerators can be used to solve problems of this type.
-Moreover, an accelerator can solve several such problems at once \cite{Kindratenko2014}; i.e., using the GPU, one can calculate multiple function values simultaneously.
+Moreover, an accelerator can solve several such problems at once \cite{Ref8}; i.e., using the GPU, one can calculate multiple function values simultaneously.
 
@@
 \section{Multidimensional parallel global search algorithm}
+Consider the problem of finding the global minimum of the \(N\)-dimensional function \(\phi(y)\) in the hyperinterval  \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\)  We will assume that the function satisfies the Lipschitz condition with an a priori unknown constant \(L\).
+\begin{equation}
+\label{task}
+\phi(y^*)=\min\{\phi(y):y\in D\}
+\end{equation}
+\begin{equation}
+\label{lip}
+|\phi(y_1)-\phi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
+\end{equation}
 
-\section{Parallel global search algorithm}
+In this paper, we will use an approach based on the idea of reducing dimensionality using the Peano space-filling curve \(y(x)\), which continuously and unambiguously maps a segment of the real axis \([0,1]\) onto an \(n\)-dimensional cube
+\begin{equation}
+\label{cube}
+\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
+\end{equation}
+
+The questions of numerical construction of approximations to Piano curve (evolvents) and the corresponding theory are discussed in detail in \cite{Ref9}, \cite{Ref10}.
+Using evolvents \(y(x)\) reduces the multidimensional problem (\ref{task}) to a one-dimensional problem
+\begin{displaymath}
+\label{oneDimTask}
+\phi(y^*)=\phi(y(x^*))=\min\{\phi(y(x)):x\in [0,1]\}
+\end{displaymath}
+
+An important property is that the relative differences of the function remain limited: if the function \(\phi(y)\) in the region \(D
+\) satisfies the Lipschitz condition, then the function \(\phi(y(x))\) in the interval \([0,1]\) will satisfy the uniform Holder condition
+\begin{displaymath}
+\label{holder}
+|\phi(y(x_1))-\phi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
+\end{displaymath}
+where the Holder constant \(H\) is related to the Lipschitz constant \(L\) by the ratio
+\begin{displaymath}
+H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}
+\end{displaymath}
+
+Therefore, without limiting generality, we can consider minimizing the one-dimensional function \(f(x)=\phi(y(x)), x\in[0,1]\), which satisfies the Holder condition.
+
+The considered algorithm for solving this problem involves constructing a sequence of points \(x_k\), in which the values of objective function \(z_k = f(x_k)\) are calculated.
+We will call the process of computing the value of a function at a single point a trial.
+Assume that we have \(P\geqslant 1\) computational elements at our disposal and \(P\) tests which are performed simultaneously (synchronously) within a single iteration of the method.
+Let \(k(n)\) denote the total number of tests performed, after \(n\) parallel iterations.
+
+In the first iteration of the method, the test is carried out at an arbitrary internal point \(x^1\) of the interval \([0,1]\).
+Let \(n>1\) iterations of the method be performed, during which tests were carried out at \(k = k(n)\) points \(x^i, 1\leqslant i\leqslant k\).
+Then the points  \(x^{k+1},\dotsc,x^{k+p}\).
+search tests of the following \((n+1)\)-th iteration are determined in accordance with the rules:
+
+Step 1.
+Renumber the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which includes the boundary points of the interval \([0,1]\), as well as the points of the previous tests, with the lower indices in the order of their increasing coordinate values, i.e.
+\begin{displaymath}
+0=x_0<x_1<\dotsc<x_{k+1}=1
+\end{displaymath}
+
+Step 2. Assuming \(z_i=f(y(x_i)),1\leqslant i\leqslant k\) calculate the values
+\begin{displaymath}
+\label{step2}
+\mu=\max_{1\leqslant i\leqslant k}\dfrac{|z_i-z_{i-1}|}{\Delta_i},
+\begin{matrix}
+    M = 
+    \left\{
+    \begin{matrix}
+    r\mu,\mu>0 \\
+    1,\mu=0
+    \end{matrix} \right.
+    \end{matrix}
+\end{displaymath}
+where \(r > 1\) is the specified parameter of the method, and \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
+
+Step 3. For each interval \((x_{i-1},x_i),1\leqslant i\leqslant k+1\), calculate the characteristic in accordance with the formulas 
+\begin{displaymath}
+\label{step3_1}
+R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M}
+\end{displaymath}
+\begin{displaymath}
+\label{step3_2}
+R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1
+\end{displaymath}
+
+Step 4. The characteristics \(R(i),1\leqslant i\leqslant k+1\), arrange in descending order 
+\begin{displaymath}
+\label{step4}
+R(t_1)\geqslant R(t_2)\geqslant \dots \geqslant R(t_{k})\geqslant R(t_{k+1})
+\end{displaymath}
+and select \(P\) of the largest characteristics with interval numbers \(t_j,1\leqslant j\leqslant P\).
+
+Step 5. Carry out new tests at the points \(x_{k+j},1\leqslant j\leqslant P\), calculated using the formulas
+\begin{displaymath}
+x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},t_j=1,t_j=k+1
+\end{displaymath}
+\begin{displaymath}
+\label{step5}
+{{x}^{k+1}}=\frac{{{x}_{{{t}_{j}}}}+{{x}_{{{t}_{j}}-1}}}{2}-\text{sign}({{z}_{{{t}_{j}}}}-{{z}_{{{t}_{j}}-1}})\frac{1}{2r}{{\left[ \frac{|{{z}_{{{t}_{j}}}}-{{z}_{{{t}_{j}}-1}}|}{\mu } \right]}^{N}}, 1<{{t}_{j}}<k+1
+\end{displaymath}
+
+The algorithm stops working if the condition \(\Delta_{t_j}\leqslant \varepsilon\)  is satisfied for at least one number \(t_j,1\leqslant j\leqslant P\); here \(\varepsilon>0\)  is the specified accuracy.
+As an estimate of the globally optimal solution to the problem (\ref{task}), the values are selected
+\begin{displaymath}
+f_{k}^{*}=\underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}}),   x_{k}^{*}=\arg \quad \underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}})
+\end{displaymath}
+
+For the rationale in using this method of organizing parallel computing, see \cite{Ref11}.
+Modifications that account for the presence of inequality constraints in the problem, as well as information about the derivative of the objective function, are presented in \cite{Ref12} – \cite{Ref14}.
 
@@
 
+Reducing multidimensional problems to one-dimensional ones through the use of evolvents has important properties such as continuity and preserving uniform bounding of function differences with limited argument variation.
+However, some information about the proximity of points in multidimensional space is lost, since the point \(x\in [0,1]\) has only left and right neighbors, and the corresponding point \(y(x) \in R^N\) has neighbors in \(2^N\) directions.
+As a result, when using evolvents, the images \(y' , y''\) that are close in \(N\)-dimensional space can correspond to rather distant prototypes \(x' , x''\) on the interval \([0,1]\).
+This property leads to redundant calculations, because several limit points \(x' , x''\) of the test sequence generated by the method on the segment \([0,1]\), can correspond to a single limit point \(y\) in \(N\)-dimensional space.
+
+A possible way to overcome this disadvantage is to use a set of evolvents (multiple mappings)
+
+\begin{displaymath}
+\label{eq:142}
+Y_L(x)=\left\{y^0(x),\ y^1(x),...,\ y^L(x)\right\}
+\end{displaymath}
+
+instead of using a single Peano curve \(y(x)\) (see \cite{Ref19}, \cite{Ref20}).
+For example, each Peano curve \(y^i(x)\) from \(Y_L(x)\) can be obtained as a result of some shift along the main diagonal of the hyperinterval \(D\), the original curve \(y(x)\).
+Another way is to rotate the sweep around the coordinate.
+The set of evolvents that have been constructed allows us to obtain for any close images \(y', y''\), close prototypes \(x', x''\) for some mapping \(y^i(x)\).
+
+   Using a set of mappings leads to the formation of a corresponding set of one-dimensional multiextremal problems
+
+\begin{displaymath}
+\label{6_problem_l}
+\min{\left\{\varphi(y^l(x)):x\in [0,1], \;  \right\}}, \ 0 \leq l \leq L.
+\end{displaymath}
+
+Each problem from this set can be solved independently, and any calculated value \(z= \varphi(y')\), \(y'=y^i(x')\) of the function \(\varphi(y)\) in the \(i\)-th problem can be interpreted as calculating the value \(z= \varphi(y')\), \(y'=y^s(x'')\) for any other \(s\)-th problem without repeated labor-intensive calculations of the function \(\varphi(y)\).
+Such informational unity makes it possible to solve the entire set of problems in a parallel fashion.
+This approach was discussed in detail in \cite{Ref21}.
+
+\subsection{Recursive dimensionality reduction scheme}
+
+The recursive optimization scheme is based on the well-known relation \cite{Ref22}
+
+\begin{equation}
+\label{nestedScheme}
+\min{\phi(y):y\in D}=\min_{a_1\leqslant y_1\leqslant b_1}\min_{a_2\leqslant y_2\leqslant b_2}\dots\min_{a_1\leqslant y_N\leqslant b_N}\phi(y)
+\end{equation}
+
+which allows one to replace the solution of the multidimensional problem (\ref{task}) with the solution of a family of one-dimensional subproblems recursively related to each other.
+Let's introduce a set of functions
+
+\begin{equation}
+\label{phiN}
+\phi_N(y_1,\dots,y_N)=\phi(y_1,\dots,y_N)
+\end{equation}
+\begin{equation}
+\label{phiI_}
+\phi_i(y_1,\dots,y_i)=\min_{a_{i+1}\leqslant y_{i+1} \leqslant b_{i+1}}\phi_{i+1}(y_1,\dots,y_i,y_{i+1}),1\leqslant i\leqslant N-1
+\end{equation}
+
+Then, in accordance with the relation (\ref{nestedScheme}) , the solution of the original problem (\ref{task}) is reduced to the solution of a one-dimensional problem
+
+\begin{equation}
+\label{phiFirst}
+\phi_1(y_1^*)=\min\{\phi_1(y_1),y_1\in [a,b]\}
+\end{equation}
+
+However, each calculation of the value of the one-dimensional function  \(\phi_1(y_1)\) at some fixed point corresponds to the solution of a one-dimensional minimization problem
+
+\begin{displaymath}
+\label{phi2}
+\phi_2(y_1,y_2^*)=\min\{\phi(y_1,y_2):y_2\in [a_2,b_2]\}
+\end{displaymath}
+
+And so on, until the calculation   according to (\ref{phiN}).
+
+For the recursive scheme described above, a generalization (block recursive scheme) is proposed that combines the use of sweeps and a recursive scheme in order to efficiently parallelize computations.
+
+Consider the vector y as a vector of block variables
+\begin{displaymath}
+y=(y_1,\dots,y_N)=(u_1,u_2,\dots,u_M)
+\end{displaymath}
+
+Where the \(i\)-th block variable \(u_i\) is a dimensional vector   of sequentially taken components of the vector \(y\), i.e.
+\(u_1=(y_1,y_2,\dots,y_{N_1}),u_2=(y_{N_1+1},y_{N_1+2},\dots,y_{N_1+N_2}),\dots,u_M=(y_{N-N_M+1},y_{N-N_M+2},\dots,y_N)\), while \(N_1+N_2+\dots+N_M=N\).
+
+Using new variables, the main relation of the multistep scheme (\ref{nestedScheme}) can be rewritten as
+
+\begin{equation}
+\label{blockNested}
+\min\phi(y)_{y\in D}=\min_{u_1\in D_1}\min_{u_2\in D_2}\dots\min_{u_M\in D_M}\phi(y)
+\end{equation}
+
+where the subdomains  \(D_i,1\leqslant i\leqslant M\), are projections of the original search domain \(D\) onto the subspaces corresponding to the variables \(u_1,1\leqslant i\leqslant M\) .
+
+The formulas that determine the method for solving problem (\ref{task}) based on relations (\ref{blockNested}) generally coincide with the recursive scheme (\ref{phiN})-(\ref{phiFirst}).
+One need only replace the original variables  \(y_i,1\leqslant i\leqslant N\), with block variables  \(u_1,1\leqslant i\leqslant M\).
+In this case, the fundamental difference from the original scheme is the fact that the block diagram has nested subproblems
+
+\begin{equation}
+\label{subTasks}
+\phi_i(u_1,\dots,u_i)=\min_{u_{i+1}\in D_{i+1}}\phi_{i+1}(u_1,\dots,u_i,u_{i+1}),1\leqslant i\leqslant M-1
+\end{equation}
+
+which are multidimensional, and to solve them, a method of reducing dimensionality based on Peano curves can be applied.
+
 \subsection{Adaptive dimensionality reduction scheme}
+
+The solution to the resulting set of subproblems (\ref{subTasks}) can be organized in various ways.
+
+The obvious method (elaborated in detail in \cite{Ref23},\cite{Ref24},\cite{Ref25} for the nested optimization scheme and in \cite{Ref26},\cite{Ref27} for the block nested optimization scheme) is based on solving subproblems in accordance with the recursive order of their generation.
+However, there is a significant loss of information about the target function.
+
+Another approach is an adaptive scheme in which all subproblems are solved simultaneously, which makes it possible to more fully take into account information about a multidimensional problem and thereby to speed up the process of solving it.
+
+In the case of one-dimensional subproblems, this approach was theoretically substantiated and tested in \cite{Ref28}, \cite{Ref29}, \cite{Ref30}, and in the article [] a generalization of the adaptive scheme for multidimensional subproblems was proposed.
+
+   The adaptive dimensionality reduction scheme changes the order in which subproblems are solved: they will be solved not one by one (in accordance with their hierarchy in the task tree), but simultaneously, i.e., there will be a number of subproblems that are in the process of being solved.
+Under the new scheme:
+
+•	to calculate the value of the function of the \(i\)-th level from (\ref{subTasks}), a new level \(i+1\) problem is generated in which tests are carried out, after which a new generated problem is included in the set of existing problems to be solved;
+
+•	the iteration of the global search consists in choosing \(P\) (the most promising) problem from the set of existing problems in which tests are carried out; points for new tests are determined in accordance with the algorithm of subsection 1;
+
+•	the minimum values of functions from (\ref{subTasks}) are their current estimates based on the accumulated search information. .
+
+A brief description of the main steps of a block adaptive dimensionality reduction scheme is as follows.
+
+Let the nested subproblems in the form (\ref{subTasks}) be solved using the global search algorithm described in subsection 1.
+Then each subproblem (\ref{subTasks}) can be assigned a numerical value called the characteristic of this problem.
+As such, we can take the maximum characteristic of the intervals formed in this problem – \(R(t)\).
+In accordance with the rule for calculating characteristics, the higher the value of the characteristic, the more promising the subproblem is in the continued search for the global minimum of the original problem (\ref{task}).
+Therefore, at each iteration, subproblems with the maximum characteristic are selected for conducting the next test.
+The test either leads to the calculation of the value of the objective function \(\phi_1(y)\) (if the selected subproblem belonged to the level \(j=M\)), or generates new subproblems according to (\ref{subTasks}) for \(j \geqslant  M-1\).
+In the latter case, the newly generated problems are added to the current set of problems, their characteristics are calculated, and the process is repeated.
+The optimization process is completed when the root problem satisfies the condition for stopping the algorithm that solves this problem.
+Some results pointing in this direction are presented in \cite{Ref31}, \cite{Ref32}.
 
@@
 
+In relation to global optimization methods, an operation that can be effectively implemented on the GPU is the parallel calculation of many values of the objective function at once.
+Naturally, this requires implementing a procedure for calculating the value of a function on the GPU.
+Data transfers from the CPU to the GPU will be minimal: one need only transfer the coordinates of the test points to the GPU, and get back the function values at these points.
+Functions that determine the processing of test results in accordance with the algorithm, and require work with large amounts of accumulated search information, can be effectively implemented on the CPU.
+
+   The general scheme for organizing calculations using the GPU is shown in \ref{fig:GPU_Algorithm}.
+In accordance with this scheme, steps 1-4 of the parallel global search algorithm are performed on the CPU.
+The coordinates of the P test points calculated in step 4 of the algorithm are accumulated in the intermediate buffer and then transmitted to the GPU.
+On the GPU, the function values are calculated at these points, after which the test results (again through the intermediate buffer) are transferred to the CPU.
+
+   \begin{figure}[h!]
+    \centering
+		\includegraphics[width=1.0\textwidth]{GPU_Algorithm.eps}
+		\caption{Scheme of information exchanges in the GPU algorithm}\label{fig:GPU_Algorithm}
+\end{figure}
+
 \subsection{Organization of parallel computing}
 
+To organize parallel calculations, we will use a set of evolvents and a block adaptive dimensionality reduction scheme.
+We take a small number of nesting levels, 2 to 3, in which the original large dimensionality problem is divided into 2-3 nested lower dimensional subproblems.
+We will use multiple sweeps only at the upper level of nesting, corresponding to the u1 variable block.
+This subproblem will be reduced to a set of one-dimensional problems that will be solved in parallel, each in a separate process.
+Test results at point x obtained for the problem solved by a specific processor are interpreted as the test results in the remaining problems (at the corresponding points \(u^1_1,… , u^s_1\) )
+Then, applying an adaptive scheme to solve the nested subproblems (\ref{subTasks}), we get a parallel algorithm with a wide degree of variability.
+
+\ref{fig:Parallel_Computing} shows the general scheme of organizing calculations using several cluster nodes and several GPUs.
+In accordance with this scheme, nested subproblems  ${{\phi }_{i}}({{u}_{1}},...,{{u}_{i}})=\underset{{{u}_{i+1}}\in {{D}_{i+1}}}{\mathop{\min }}\,{{\phi }_{i+1}}({{u}_{1}},...,{{u}_{i}},{{u}_{i+1}})$  with \(i=1,\ldots ,M-2\) are solved using only the CPU.
+The values of the optimized function are not calculated directly in these subproblems: the calculation the function value  ${{\phi }_{i}}({{u}_{1}},...,{{u}_{i}})$ is a solution to the minimization problem at the next level.
+
+The subproblem of the last \((M - 1)\) level \({{\phi }_{M-1}}({{u}_{1}},...,{{u}_{M-1}})=\underset{{{u}_{M}}\in {{D}_{M}}}{\mathop{\min }}\,{{\phi }_{M}}({{u}_{1}},...,{{u}_{M}})\) differs from all the previous subproblems – it calculates the values of the optimized function, since ${{\phi }_{M}}({{u}_{1}},...,{{u}_{M}})=\phi ({{y}_{1}},...,{{y}_{N}})$.
+This subproblem transfers between the CPU and the GPU.
+
+   \begin{figure}[h!]
+    \centering
+		\includegraphics[width=1.0\textwidth]{Parallel_Computing.eps}
+		\caption{Diagram of parallel computing on a cluster}\label{fig:Parallel_Computing}
+\end{figure}
+
 \section{Numerical experiments}
+
+The numerical experiments were carried out using the Lomonosov supercomputer (Lomonosov Moscow State University).
+Each supercomputer node included two quad-core processors, Intel Xeon X5570 2.93GHz and 12 Gb RAM.
+When building the Globalizer system for running on the Lomonosov supercomputer, the GCC 4.3.0 compiler, CUDA 6.5 and Intel MPI 2017 were used.
+
+Note that well-known test problems from the field of multidimensional global optimization are characterized by a short time span for calculating the values of the objective function.
+Therefore, in order to simulate the computational complexity inherent in applied optimization problems, the calculation of the objective function in all experiments was complicated by additional calculations that do not change the form of the function and the location of its minima (summing the segment of the Taylor series).
+
+In the study \cite{Ref15}, \cite{Ref16}, a GKLS generator is described that allows one to generate multiextremal optimization problems with previously known properties: the number of local minima, the size of their regions of attraction, the global minimum point, the value of the function in it, etc.
+
+Below are the results of a numerical comparison of three sequential algorithms -- DIRECT \cite{Ref15}, DIRECTl \cite{Ref17} and Global Search Algorithm (GSA); (results of the first two algorithms are given in the article \cite{Ref18}).
+A numerical comparison was carried out on the Simple and Hard function classes of dimension 4 and 5 from \cite{Ref18}.
+The Global minimum $y*$ was considered found if the algorithm generated a test point $y^k$ in the $\delta$-neighborhood of the global minimum, i.e.
+$\left\| {{y}^{k}}-{{y}^{*}} \right\|\le \delta $.
+Also, the size of the neighborhood was chosen (in accordance with \cite{Ref18}) as $\delta =\left\| b-a \right\|\sqrt[N]{\Delta }$, here $N$ is the dimension of the problem to be solved, a и b are the boundaries of the search domain $D$, the parameter  $\Delta $=10-6 for N=4 and $\Delta =10-7$ for $N=5$.
+When using the AGP method for the Simple class, the parameter $r=4.5$ was selected, for the Hard class - $r=5.6$; the parameter for constructing the Peano curve was fixed, $m=10$.
+The maximum number of iterations allowed was $Kmax = 1 000 000$.
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|p{3cm}|c|c|c|}
+    \hline
+    \(N\) & Problem class & DIRECT & DIRECT\(l\) & GSA \\ \hline
+    4 & Simple & >47282 (4) & 18983 & 11953\\ \cline{2-5}
+      & Hard & >95708 (7) & 68754 & 25263\\ \hline
+    5 & Simple & >16057 (1) & 16758 & 15920\\ \cline{2-5}
+      & Hard & >217215 (16) & >269064 (4) & >148342 (4)\\ \hline
+    \end{tabular}
+    \caption{Average number of iterations}
+    \label{table:average_iters}
+\end{table}
+
+   Table  \ref{table:average_iters} shows the average number of iterations, $kav$, that the method performed when solving a series of problems from these classes.
+The symbol ">" reflects a situation where not all problems of the class were solved by any method whatsoever.
+This means that the algorithm was halted because the maximum allowed number of $Kmax$ iterations was reached.
+In this case, the value $Kmax = 1 000 000$ was used to calculate the average value of the number of iterations, $kav$, which corresponds to the lower estimate of this average value.
+The number of unsolved problems is indicated in parentheses.
+
+  As can be seen from Table \ref{table:average_iters}, the sequential GSA surpasses the DIRECT and DIRECTl methods in all classes of problems in terms of the average number of iterations.
+At the same time, in the 5-Hard class, none of the methods solved all the problems: DIRECT failed to solve 16 problems, DIRECTl and GSA -- 4 problems each.
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|p{3cm}|c|c|c|}
+    \hline
+    \(N\) & Problem class & P=2 & P=4 & P=8 \\ \hline
+    4 & Simple  & 3,49 & 8,22  &  \\ \cline{2-5}
+      & Hard    & 2,35 & 7,65  &  \\ \hline
+    5 & Simple  & 1,87 & 6,59  &  \\ \cline{2-5}
+      & Hard    & 1,48 & 10,80  &   \\ \hline
+    \end{tabular}
+    \caption{Acceleration of iterations on the CPU}
+    \label{table:average_iters_AS_CPU}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|p{3cm}|c|c|c|}
+    \hline
+    \(N\) & Problem class & P=2 & P=4 & P=8 \\ \hline
+    4 & Simple  & 1,71 & 2,01  &  \\ \cline{2-5}
+      & Hard    & 1,17 & 1,90  &  \\ \hline
+    5 & Simple  & 0,88 & 1,48  &  \\ \cline{2-5}
+      & Hard    & 0,70 & 2,39  &   \\ \hline
+    \end{tabular}
+    \caption{Time acceleration on the CPU}
+    \label{table:average_Time_AS_CPU}
+\end{table}
+
+   Let us now evaluate the acceleration achieved using parallel GSA using an adaptive dimensionality reduction scheme based on the number $P$ of cores used.
+Table \ref{table:average_iters_AS_CPU} shows the acceleration by iterations, and Table \ref{table:average_Time_AS_CPU} shows the time acceleration obtained when solving a series of problems on the CPU, compared to GSA.
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|p{3cm}|c|c|c|}
+    \hline
+    \(N\) & Problem class & P=2 & P=4 & P=8 \\ \hline
+    4 & Simple &   6,62 & 21,46  &  \\ \cline{2-5}
+      & Hard   &   3,23 & 9,97  &  \\ \hline
+    5 & Simple &   3,37 & 11,20  &  \\ \cline{2-5}
+      & Hard   &   2,52 & 8,54  &   \\ \hline
+    \end{tabular}
+    \caption{Time acceleration on the CPU}
+    \label{table:average_iters_MMAS_CPU}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|p{3cm}|c|c|c|}
+    \hline
+    \(N\) & Problem class & P=2 & P=4 & P=8 \\ \hline
+    4 & Simple &   3,22 & 5,23  &  \\ \cline{2-5}
+      & Hard   &   1,48 & 2,42  &  \\ \hline
+    5 & Simple &   1,30 & 1,97  &  \\ \cline{2-5}
+      & Hard   &   0,68 & 0,88  &   \\ \hline
+    \end{tabular}
+    \caption{Acceleration of iterations on the CPU}
+    \label{table:average_Time_MMAS_CPU}
+\end{table}
+
+   Table \ref{table:average_iters_MMAS_CPU} shows the acceleration in iterations, and Table \ref{table:average_Time_MMAS_CPU} shows the acceleration in time of the algorithm that combines multiple sweeps and an adaptive scheme for solving a series of problems on the CPU, compared to the sequential launch of the GSA.
+Two scans and, accordingly, two processes were used; each process used $P$ threads, and calculations were performed on a single cluster node.
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|p{3cm}|c|c|c|}
+    \hline
+    \(N\) & Problem class & P=64 & P=256 & P=1024 \\ \hline
+    4 & Simple &   5,06 & 19,91 & 15,62  \\ \cline{2-5}
+      & Hard   &   3,93 & 14,95 &   \\ \hline
+    5 & Simple &   4,16 & 16,12 & 67,22  \\ \cline{2-5}
+      & Hard   &   3,36 & 10,02 & 42,04   \\ \hline
+    \end{tabular}
+    \caption{Acceleration of iterations on the GPU}
+    \label{table:average_iters_AS_GPU}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|p{3cm}|c|c|c|}
+    \hline
+    \(N\) & Problem class & P=64 & P=256 & P=1024 \\ \hline
+    4 & Simple &   2,20 & 8,28 & 5,21  \\ \cline{2-5}
+      & Hard   &   1,92 & 6,91 &   \\ \hline
+    5 & Simple &   1,58 & 4,18 & 8,23  \\ \cline{2-5}
+      & Hard   &   1,22 & 2,39 & 3,74   \\ \hline
+    \end{tabular}
+    \caption{Time acceleration on the GPU}
+    \label{table:average_Time_AS_GPU}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|p{3cm}|c|c|c|}
+    \hline
+    \(N\) & Problem class & P=64 & P=256 & P=1024 \\ \hline
+    4 & Simple &   11,32 & 38,96 & 128,39  \\ \cline{2-5}
+      & Hard   &   6,83  & 31,52 &   \\ \hline
+    5 & Simple &   4,99  & 39,37 & 80,56  \\ \cline{2-5}
+      & Hard   &   4,61  & 17,28 & 113,61   \\ \hline
+    \end{tabular}
+    \caption{Acceleration of iterations on the GPU}
+    \label{table:average_iters_MMAS_GPU}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|p{3cm}|c|c|c|}
+    \hline
+    \(N\) & Problem class & P=64 & P=256 & P=1024 \\ \hline
+    4 & Simple &   2,44 & 8,01 & 20,45  \\ \cline{2-5}
+      & Hard   &   1,63 & 7,04 &   \\ \hline
+    5 & Simple &   0,77 & 3,97 & 2,37  \\ \cline{2-5}
+      & Hard   &   0,61 & 0,96 & 2,11   \\ \hline
+    \end{tabular}
+    \caption{Time acceleration on the GPU}
+    \label{table:average_Time_MMAS_GPU}
+\end{table}
+
+Tables \ref{table:average_iters_AS_GPU} and  \ref{table:average_Time_AS_GPU} show the acceleration obtained when solving a series of problems on the GPU using an adaptive scheme compared to a similar launch on the CPU using 4 threads.
+Tables \ref{table:average_iters_MMAS_GPU} and \ref{table:average_Time_MMAS_GPU} show the acceleration of the algorithm combining multiple sweeps and an adaptive circuit when solving a series of problems on the GPU, compared to an adaptive scheme on the CPU using 4 threads.
+Two sweeps and, accordingly, two processes were used; each process used P threads, and calculations were performed on a single cluster node.
+
+\begin{table}
+    \centering
+    \begin{tabular}{|p{4cm}|c|c|c|}
+    \hline
+      & P=2 & P=4 & P=8 \\ \hline
+    Acceleration of iterations&   30,79 & 92,68 & 596,97  \\ \hline
+    Time acceleration         &   1,86  & 1,49  & 2,45    \\ \hline
+
+    \end{tabular}
+    \caption{Acceleration of iterations and time on the GPU}
+    \label{table:GKLS6_average_MMAS_GPU}
+\end{table}
+
+Следующий эксперимент будет проведен на 6-ти мерных задачах GKLS.
+В таблице \ref{table:GKLS6_average_MMAS_GPU} приведено ускорение по времени и по итерациям полученное при решении серии задач на GPU с использованием алгоритма объединяющего множественные развертки и адаптивную схему по сравнению с адаптивной схемой на CPU использующей 4 потока.
+Эксперимент проводился на трех узлах кластера и спользовалось до 6144 нитей.
 
@@
 %
-\bibliographystyle{spmpsci}
-\bibliography{bibliography}{}
+\begin{thebibliography}{6}
+
+\bibitem {Ref1}	Hwu, W. GPU Computing Gems: Emerald Edition — San Francisco: Morgan Kaufmann, 2011. — 886 p.
+\bibitem {Ref2}	A.M.
+Ferreiro, J.A.
+Garcia, J.G.
+Lopez-Salas, C.
+Vazquez.
+An efficient implementation of parallel simulated annealing algorithm in GPUs // Journal of global optimization.
+— 2013.
+— V.
+57, No 3.
+— P.
+863–890.
+\bibitem {Ref3}	J.M. Garcia-Martinez, E.M. Garzon, P.M. Ortigosa. A GPU implementation of a hybrid evolutionary algorithm: GPuEGO // Journal of super-computing. — 2014. — V. 70, No 2. — P. 684–695.
+\bibitem {Ref4}	Langdon, W.B. Graphics processing units and genetic programming: an overview // Soft Computing. — 2011. — V. 15, No 8. — P. 1657–1669.
+\bibitem {Ref5}	Ю.Г.
+Евтушенко, В.У.
+Малкова, А.А.
+Станевичюс.
+Параллельный поиск глобального экстремума функций многих переменных // Ж.
+вычисл.
+матем.
+и матем.
+физ.
+— 2009.
+— Т.
+49, No2.
+— С.
+255–269.
+\bibitem {Ref6}	J.
+He, A.
+Verstak, L.T.
+Watson, M.
+Sosonkina.
+Design and implementation of a massively parallel version of DIRECT // Computational optimization and applications.
+— 2008.
+— V.
+40, No 2.
+— P.
+217–245.
+\bibitem {Ref7}	R.
+Paulavicius, J.
+Žilinskas, A.
+Grothey.
+Parallel branch and bound for global optimization with combination of Lipschitz bounds // Optimization methods and software.
+— 2011.
+— V.
+26, No 3.
+— P.
+487–498.
+\bibitem {Ref8}	V. Kindratenko (ed.). Numerical computations with GPUs — New York: Springer, 2014. — 405 p.
+\bibitem {Ref9}	Strongin, R.G. Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms / R.G. Strongin, Ya.D. Sergeyev – Kluwer Academic Publishers, 2000. – 704 p.
+\bibitem {Ref10}	Sergeyev, Ya.D. Introduction to global optimization exploiting space-filling curves / Ya.D. Sergeyev, R.G. Strongin, D. Lera – Springer, 2013. – 125 p.
+\bibitem {Ref11}	Стронгин, Р.Г.
+Параллельные вычисления в задачах глобальной оптимизации / Р.Г.
+Стронгин, В.П.
+Гергель, В.А.
+Гришагин, К.А.
+Баркалов – М.: Издательство Московского университета, 2013.
+– 280 с.
+\bibitem {Ref12}	Barkalov, K.A.
+A global optimization technique with an adaptive order of checking for constraints / K.A.
+Barkalov, R.G.
+Strongin // Computational Mathematics and Mathematical Physics.
+– 2002.
+– Vol.
+42, No.
+9.
+– P.
+1289-1300.
+\bibitem {Ref13}	Gergel, V.P.
+A method of using derivatives in the minimization of multiextremum functions / V.P.
+Gergel // Computational Mathematics and Mathematical Physics.
+– 1996.
+– Vol.
+36, No.
+6.
+– P.
+729 – 742.
+\bibitem {Ref14}	Gergel, V.P.
+A global optimization algorithm for multivariate functions with lipschitzian first derivatives / V.P.
+Gergel // Journal of Global Optimization.
+– 1997.
+– Vol.
+10, No.
+3.
+– P.
+257-281.
+\bibitem {Ref15}	Gaviano, M.
+Software for generation of classes of test functions with known local and global minima for global optimization/ M.
+Gaviano, D.
+Lera, D.
+E.
+Kvasov, Y.
+D.
+Sergeyev // ACM Transactions on Mathematical Software.
+– 2003.
+– Vol.
+29.
+– P.
+469-480.
+\bibitem {Ref16}	Сергеев, Я.Д. Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. – 352 c.
+\bibitem {Ref17}	Gablonsky, J.M. A Locally-Biased Form of the DIRECT Algorithm / J.M.  Gablonsky, C.T. Kelley, // Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.
+\bibitem {Ref18}	Sergeyev, Ya.D.
+Global search based on efficient diagonal partitions and a set of Lipschitz constants / Ya.D.
+Sergeyev, D.E.
+Kvasov // SIAM Journal on Optimization.
+– 2006.
+Vol.
+16, No.
+3.
+– P.
+910–937.
+\bibitem {Ref19}	Strongin R.G. Algorithms for multi-extremal mathematical programming problems employ-ing the set of joint space-filling curves// J. of Global Optimization, 1992. No2. P. 357–378. 
+\bibitem {Ref20}	 Strongin R.G., Sergeyev Ya.D. Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht, 2000. 
+\bibitem {Ref21}	Barkalov K., Lebedev I., Sovrasov V.
+Comparison of dimensionality reduction schemes for parallel global optimization algorithms.
+Communications in Computer and Information Science Volume 965, 2019, Pages 50-62
+\bibitem {Ref22}	Городецкий С.Ю., Гришагин В.А. Нелинейное программирование и многоэкстремальная оптимизация. Н.Новгород: Изд-во ННГУ, 2007.
+\bibitem {Ref23}	Grishagin V.A., Sergeyev Y.D., Strongin R.G. Parallel Characteristical Algorithms for Solving Problems of Global Optimization // J. Glob. Optim. 1997. V.~10~(2). P. 185--206.
+\bibitem {Ref24}	Sergeyev Y., Grishagin V. Parallel Asynchronous Global Search and the Nested Optimization Scheme // J. Comput. Anal. Appl. 2001. V. 3~(2). P. 123--145.
+\bibitem {Ref25}	Gergel V., Grishagin V., Israfilov R. Local Tuning in Nested Scheme of Global Optimization // Procedia Computer Sci. 2015. V.~51~(1). P.~865--874.
+\bibitem {Ref26}	Barkalov K., Gergel V.
+Multilevel Scheme of Dimensionality Reduction for Parallel Global Search Algorithms / OPT-i 2014 --- 1st Int.
+Conf.
+on Engineering and Applied Sciences Optimization.
+Proc.
+2014.
+P.
+2111--2124.
+\bibitem {Ref27}	Barkalov K., Lebedev I. Solving Multidimensional Global Optimization Problems using Graphics Accelerators // CCIS. 2016. V. 687. P. 224--235.
+\bibitem {Ref28}	Gergel V., Grishagin V., Gergel A. Adaptive Nested Optimization Scheme for Multidimensional Global Search // J. Glob. Optim. 2016. V.~66~(1). P.~35--51.
+\bibitem {Ref29}	Grishagin V., Israfilov R., Sergeyev Y. Comparative Efficiency of Dimensionality Reduction Schemes in Global Optimization // AIP Conf. Proc. 2016. V. 1776. Art. No. 060011.
+\bibitem {Ref30}	Grishagin V., Israfiov R., Sergeyev Y.
+Convergence Conditions and Numerical Comparison of Global Optimization Methods Based on Dimensionality Reduction Schemes // Appl.
+Math.
+Comput.
+2018.
+V.
+318.
+P.
+270--280.
+\bibitem {Ref31}	Gergel V., Grishagin V., Israfilov R. Parallel Dimensionality Reduction for Multiextremal Optimization Problems // LNCS. 2019. V. 11657. P. 166--178.
+\bibitem {Ref32}	Стронгин Р.Г., Гергель В.П., Баркалов К.А. Адаптивная глобальная оптимизация на основе блочно-рекурсивной схемы редукции размерности.
+
+\end{thebibliography}
 

### 2020-04-13 2020/RSD_GPU/RSD2020.tex
Fix some mistakes

@@
 \usepackage{amssymb}
-\usepackage[russian]{babel}
+
+%\usepackage[russian]{babel}
 
@@
 %
-\author{Konstantin Barkalov\letter\orcidID{0000-0001-5273-2471} \and \\  Ilya Lebedev\orcidID{0000-0002-8736-0652}}
+\author{Konstantin Barkalov\letter%\orcidID{0000-0001-5273-2471} 
+\and Ilya Lebedev%\orcidID{0000-0002-8736-0652}
+}
+
 %
@@
 
-%\maketitle              % typeset the title of the contribution
+\maketitle              % typeset the title of the contribution
 
@@
 Additionally, in the past few years some principles for developing graphics hardware have changed, and as a result it has become more programmable.
-Today, a graphics accelerator is a flexibly programmable, massive parallel processor with high performance, which is in demand for solving a range of computationally time-consuming problems \cite{Ref1}.
+Today, a graphics accelerator is a flexibly programmable, massive parallel processor with high performance, which is in demand for solving a range of computationally time-consuming problems \cite{Hwu2011}.
 
 However, the potential for graphics accelerators to solve global optimization problems has not yet been fully realized.
-Using the GPUs, they basically parallelize nature-inspired optimization algorithms, which are somehow based on the idea of random search (see, for example, \cite{Ref2,Ref3,Ref4}).
+Using the GPUs, they basically parallelize nature-inspired optimization algorithms, which are somehow based on the idea of random search (see, for example, \cite{Ferreiro2013,Garcia2014,Langdon2011}).
 By virtue of their stochastic nature, algorithms of this type guarantee convergence to the global minimum only in the sense of probability, which differentiates them unfavorably from deterministic methods.
 
-With regard to many deterministic algorithms of Lipschitzian global optimization with guaranteed convergence, parallel variants have been proposed \cite{Ref5,Ref6,Ref7}.
+With regard to many deterministic algorithms of Lipschitzian global optimization with guaranteed convergence, parallel variants have been proposed \cite{Evtushenko2009,He2008,Paulavicius2011}.
 However, these versions of algorithms are parallelized on the CPU using shared and/or distributed memory; presently, no GPU implementations have been made.
-For example, \cite{Ref7} describes parallelization of an algorithm based on the ideas of the branch and boundary method using MPI and OpenMP.
+For example, \cite{Paulavicius2011} describes parallelization of an algorithm based on the ideas of the branch and boundary method using MPI and OpenMP.
 
@@
 Currently, graphics accelerators can be used to solve problems of this type.
-Moreover, an accelerator can solve several such problems at once \cite{Ref8}; i.e., using the GPU, one can calculate multiple function values simultaneously.
+Moreover, an accelerator can solve several such problems at once \cite{Kindratenko2014}; i.e., using the GPU, one can calculate multiple function values simultaneously.
 
@@
 \section{Multidimensional parallel global search algorithm}
-Consider the problem of finding the global minimum of the \(N\)-dimensional function \(\phi(y)\) in the hyperinterval  \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\)  We will assume that the function satisfies the Lipschitz condition with an a priori unknown constant \(L\).
+Consider the problem of finding the global minimum of the \(N\)-dimensional function \(\varphi(y)\) in the hyperinterval  \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\)  We will assume that the function satisfies the Lipschitz condition with an a priori unknown constant \(L\).
 \begin{equation}
 \label{task}
-\phi(y^*)=\min\{\phi(y):y\in D\}
+\varphi(y^*)=\min\{\varphi(y):y\in D\}
 \end{equation}
@@
 \label{lip}
-|\phi(y_1)-\phi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
+|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
 \end{equation}
@@
 
-The questions of numerical construction of approximations to Piano curve (evolvents) and the corresponding theory are discussed in detail in \cite{Ref9}, \cite{Ref10}.
+The questions of numerical construction of approximations to Piano curve (evolvents) and the corresponding theory are discussed in detail in \cite{Strongin2000}, \cite{Sergeyev2013}.
 Using evolvents \(y(x)\) reduces the multidimensional problem (\ref{task}) to a one-dimensional problem
@@
 \label{oneDimTask}
-\phi(y^*)=\phi(y(x^*))=\min\{\phi(y(x)):x\in [0,1]\}
+\varphi(y^*)=\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}
 \end{displaymath}
 
-An important property is that the relative differences of the function remain limited: if the function \(\phi(y)\) in the region \(D
-\) satisfies the Lipschitz condition, then the function \(\phi(y(x))\) in the interval \([0,1]\) will satisfy the uniform Holder condition
+An important property is that the relative differences of the function remain limited: if the function \(\varphi(y)\) in the region \(D
+\) satisfies the Lipschitz condition, then the function \(\varphi(y(x))\) in the interval \([0,1]\) will satisfy the uniform Holder condition
 \begin{displaymath}
 \label{holder}
-|\phi(y(x_1))-\phi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
+|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
 \end{displaymath}
@@
 
-Therefore, without limiting generality, we can consider minimizing the one-dimensional function \(f(x)=\phi(y(x)), x\in[0,1]\), which satisfies the Holder condition.
+Therefore, without limiting generality, we can consider minimizing the one-dimensional function \(f(x)=\varphi(y(x)), x\in[0,1]\), which satisfies the Holder condition.
 
@@
 
-instead of using a single Peano curve \(y(x)\) (see \cite{Ref19}, \cite{Ref20}).
+instead of using a single Peano curve \(y(x)\) (see \cite{Strongin1992}, \cite{Strongin2000}).
 For example, each Peano curve \(y^i(x)\) from \(Y_L(x)\) can be obtained as a result of some shift along the main diagonal of the hyperinterval \(D\), the original curve \(y(x)\).
@@
 Such informational unity makes it possible to solve the entire set of problems in a parallel fashion.
-This approach was discussed in detail in \cite{Ref21}.
+This approach was discussed in detail in \cite{Barkalov2019}.
 
@@
 The recursive optimization scheme is based on the well-known relation \cite{Ref22}
-
 \begin{equation}
 \label{nestedScheme}
-\min{\phi(y):y\in D}=\min_{a_1\leqslant y_1\leqslant b_1}\min_{a_2\leqslant y_2\leqslant b_2}\dots\min_{a_1\leqslant y_N\leqslant b_N}\phi(y)
+\min{\varphi(y):y\in D}=\min_{a_1\leqslant y_1\leqslant b_1}\min_{a_2\leqslant y_2\leqslant b_2}\dots\min_{a_1\leqslant y_N\leqslant b_N}\varphi(y)
 \end{equation}
-
 which allows one to replace the solution of the multidimensional problem (\ref{task}) with the solution of a family of one-dimensional subproblems recursively related to each other.
@@
 \label{phiN}
-\phi_N(y_1,\dots,y_N)=\phi(y_1,\dots,y_N)
+\varphi_N(y_1,\dots,y_N)=\varphi(y_1,\dots,y_N)
 \end{equation}
@@
 \label{phiI_}
-\phi_i(y_1,\dots,y_i)=\min_{a_{i+1}\leqslant y_{i+1} \leqslant b_{i+1}}\phi_{i+1}(y_1,\dots,y_i,y_{i+1}),1\leqslant i\leqslant N-1
+\varphi_i(y_1,\dots,y_i)=\min_{a_{i+1}\leqslant y_{i+1} \leqslant b_{i+1}}\varphi_{i+1}(y_1,\dots,y_i,y_{i+1}),1\leqslant i\leqslant N-1
 \end{equation}
@@
 \label{phiFirst}
-\phi_1(y_1^*)=\min\{\phi_1(y_1),y_1\in [a,b]\}
+\varphi_1(y_1^*)=\min\{\varphi_1(y_1),y_1\in [a,b]\}
 \end{equation}
 
-However, each calculation of the value of the one-dimensional function  \(\phi_1(y_1)\) at some fixed point corresponds to the solution of a one-dimensional minimization problem
+However, each calculation of the value of the one-dimensional function  \(\varphi_1(y_1)\) at some fixed point corresponds to the solution of a one-dimensional minimization problem
 
@@
 \label{phi2}
-\phi_2(y_1,y_2^*)=\min\{\phi(y_1,y_2):y_2\in [a_2,b_2]\}
+\varphi_2(y_1,y_2^*)=\min\{\varphi(y_1,y_2):y_2\in [a_2,b_2]\}
 \end{displaymath}
@@
 
-For the recursive scheme described above, a generalization (block recursive scheme) is proposed that combines the use of sweeps and a recursive scheme in order to efficiently parallelize computations.
+For the recursive scheme described above, a generalization (block recursive scheme) is proposed that combines the use of evolvents and a recursive scheme in order to efficiently parallelize computations.
 
-Consider the vector y as a vector of block variables
+Consider the vector $y$ as a vector of block variables
 \begin{displaymath}
-y=(y_1,\dots,y_N)=(u_1,u_2,\dots,u_M)
+y=(y_1,\dots,y_N)=(u_1,u_2,\dots,u_M),
 \end{displaymath}
-
-Where the \(i\)-th block variable \(u_i\) is a dimensional vector   of sequentially taken components of the vector \(y\), i.e.
-\(u_1=(y_1,y_2,\dots,y_{N_1}),u_2=(y_{N_1+1},y_{N_1+2},\dots,y_{N_1+N_2}),\dots,u_M=(y_{N-N_M+1},y_{N-N_M+2},\dots,y_N)\), while \(N_1+N_2+\dots+N_M=N\).
+where the \(i\)-th block variable \(u_i\) is a vector of sequentially taken components of the vector \(y\), i.e.
+$u_1=(y_1,y_2,\dots,y_{N_1})$,$u_2=(y_{N_1+1},y_{N_1+2},\dots,y_{N_1+N_2})$,$\dots,u_M=(y_{N-N_M+1},y_{N-N_M+2},\dots,y_N)$, while \(N_1+N_2+\dots+N_M=N\).
 
@@
 \label{blockNested}
-\min\phi(y)_{y\in D}=\min_{u_1\in D_1}\min_{u_2\in D_2}\dots\min_{u_M\in D_M}\phi(y)
+\min_{y\in D} \varphi(y)=\min_{u_1\in D_1}\min_{u_2\in D_2}\dots\min_{u_M\in D_M}\varphi(y)
 \end{equation}
@@
 \label{subTasks}
-\phi_i(u_1,\dots,u_i)=\min_{u_{i+1}\in D_{i+1}}\phi_{i+1}(u_1,\dots,u_i,u_{i+1}),1\leqslant i\leqslant M-1
+\varphi_i(u_1,\dots,u_i)=\min_{u_{i+1}\in D_{i+1}}\varphi_{i+1}(u_1,\dots,u_i,u_{i+1}),1\leqslant i\leqslant M-1
 \end{equation}
@@
 
-The obvious method (elaborated in detail in \cite{Ref23},\cite{Ref24},\cite{Ref25} for the nested optimization scheme and in \cite{Ref26},\cite{Ref27} for the block nested optimization scheme) is based on solving subproblems in accordance with the recursive order of their generation.
+The obvious method (elaborated in detail in \cite{Grishagin1997,Grishagin2001,Grishagin2015} for the nested optimization scheme and in \cite{Barkalov2014},\cite{Ref27} for the block nested optimization scheme) is based on solving subproblems in accordance with the recursive order of their generation.
 However, there is a significant loss of information about the target function.
@@
 
-In the case of one-dimensional subproblems, this approach was theoretically substantiated and tested in \cite{Ref28}, \cite{Ref29}, \cite{Ref30}, and in the article [] a generalization of the adaptive scheme for multidimensional subproblems was proposed.
+In the case of one-dimensional subproblems, this approach was theoretically substantiated and tested in \cite{Grishagin2016,Grishagin2016_1,Grishagin2018}, and in the article [] a generalization of the adaptive scheme for multidimensional subproblems was proposed.
 
@@
 Therefore, at each iteration, subproblems with the maximum characteristic are selected for conducting the next test.
-The test either leads to the calculation of the value of the objective function \(\phi_1(y)\) (if the selected subproblem belonged to the level \(j=M\)), or generates new subproblems according to (\ref{subTasks}) for \(j \geqslant  M-1\).
+The test either leads to the calculation of the value of the objective function \(\varphi_1(y)\) (if the selected subproblem belonged to the level \(j=M\)), or generates new subproblems according to (\ref{subTasks}) for \(j \geqslant  M-1\).
 In the latter case, the newly generated problems are added to the current set of problems, their characteristics are calculated, and the process is repeated.
@@
 \ref{fig:Parallel_Computing} shows the general scheme of organizing calculations using several cluster nodes and several GPUs.
-In accordance with this scheme, nested subproblems  ${{\phi }_{i}}({{u}_{1}},...,{{u}_{i}})=\underset{{{u}_{i+1}}\in {{D}_{i+1}}}{\mathop{\min }}\,{{\phi }_{i+1}}({{u}_{1}},...,{{u}_{i}},{{u}_{i+1}})$  with \(i=1,\ldots ,M-2\) are solved using only the CPU.
-The values of the optimized function are not calculated directly in these subproblems: the calculation the function value  ${{\phi }_{i}}({{u}_{1}},...,{{u}_{i}})$ is a solution to the minimization problem at the next level.
+In accordance with this scheme, nested subproblems  ${{\varphi }_{i}}({{u}_{1}},...,{{u}_{i}})=\underset{{{u}_{i+1}}\in {{D}_{i+1}}}{\mathop{\min }}\,{{\varphi }_{i+1}}({{u}_{1}},...,{{u}_{i}},{{u}_{i+1}})$  with \(i=1,\ldots ,M-2\) are solved using only the CPU.
+The values of the optimized function are not calculated directly in these subproblems: the calculation the function value  ${{\varphi }_{i}}({{u}_{1}},...,{{u}_{i}})$ is a solution to the minimization problem at the next level.
 
-The subproblem of the last \((M - 1)\) level \({{\phi }_{M-1}}({{u}_{1}},...,{{u}_{M-1}})=\underset{{{u}_{M}}\in {{D}_{M}}}{\mathop{\min }}\,{{\phi }_{M}}({{u}_{1}},...,{{u}_{M}})\) differs from all the previous subproblems – it calculates the values of the optimized function, since ${{\phi }_{M}}({{u}_{1}},...,{{u}_{M}})=\phi ({{y}_{1}},...,{{y}_{N}})$.
+The subproblem of the last \((M - 1)\) level \({{\varphi }_{M-1}}({{u}_{1}},...,{{u}_{M-1}})=\underset{{{u}_{M}}\in {{D}_{M}}}{\mathop{\min }}\,{{\varphi }_{M}}({{u}_{1}},...,{{u}_{M}})\) differs from all the previous subproblems – it calculates the values of the optimized function, since ${{\varphi }_{M}}({{u}_{1}},...,{{u}_{M}})=\varphi ({{y}_{1}},...,{{y}_{N}})$.
 This subproblem transfers between the CPU and the GPU.
@@
 
-In the study \cite{Ref15}, \cite{Ref16}, a GKLS generator is described that allows one to generate multiextremal optimization problems with previously known properties: the number of local minima, the size of their regions of attraction, the global minimum point, the value of the function in it, etc.
+In the study \cite{Gaviano2003}, \cite{Ref16}, a GKLS generator is described that allows one to generate multiextremal optimization problems with previously known properties: the number of local minima, the size of their regions of attraction, the global minimum point, the value of the function in it, etc.
 
-Below are the results of a numerical comparison of three sequential algorithms -- DIRECT \cite{Ref15}, DIRECTl \cite{Ref17} and Global Search Algorithm (GSA); (results of the first two algorithms are given in the article \cite{Ref18}).
-A numerical comparison was carried out on the Simple and Hard function classes of dimension 4 and 5 from \cite{Ref18}.
+Below are the results of a numerical comparison of three sequential algorithms -- DIRECT \cite{Jones2009}, DIRECT\textit{l} \cite{Gablonsky2001} and Global Search Algorithm (GSA); (results of the first two algorithms are given in the article \cite{Sergeyev2006}).
+A numerical comparison was carried out on the Simple and Hard function classes of dimension 4 and 5 from \cite{Sergeyev2006}.
 The Global minimum $y*$ was considered found if the algorithm generated a test point $y^k$ in the $\delta$-neighborhood of the global minimum, i.e.
 $\left\| {{y}^{k}}-{{y}^{*}} \right\|\le \delta $.
-Also, the size of the neighborhood was chosen (in accordance with \cite{Ref18}) as $\delta =\left\| b-a \right\|\sqrt[N]{\Delta }$, here $N$ is the dimension of the problem to be solved, a и b are the boundaries of the search domain $D$, the parameter  $\Delta $=10-6 for N=4 and $\Delta =10-7$ for $N=5$.
+Also, the size of the neighborhood was chosen (in accordance with \cite{Sergeyev2006}) as $\delta =\left\| b-a \right\|\sqrt[N]{\Delta }$, here $N$ is the dimension of the problem to be solved, a и b are the boundaries of the search domain $D$, the parameter  $\Delta $=10-6 for N=4 and $\Delta =10-7$ for $N=5$.
 When using the AGP method for the Simple class, the parameter $r=4.5$ was selected, for the Hard class - $r=5.6$; the parameter for constructing the Peano curve was fixed, $m=10$.
-The maximum number of iterations allowed was $Kmax = 1 000 000$.
+The maximum number of iterations allowed was $K_{max} = 10^6$.
 
@@
       & Hard & >95708 (7) & 68754 & 25263\\ \hline
-    5 & Simple & >16057 (1) & 16758 & 15920\\ \cline{2-5}
+    5 & Simple &>16057 (1) & 16758 & 15920\\ \cline{2-5}
       & Hard & >217215 (16) & >269064 (4) & >148342 (4)\\ \hline
@@
 
-   Table  \ref{table:average_iters} shows the average number of iterations, $kav$, that the method performed when solving a series of problems from these classes.
+   Table \ref{table:average_iters} shows the average number of iterations, $k_{av}$, that the method performed when solving a series of problems from these classes.
 The symbol ">" reflects a situation where not all problems of the class were solved by any method whatsoever.
-This means that the algorithm was halted because the maximum allowed number of $Kmax$ iterations was reached.
-In this case, the value $Kmax = 1 000 000$ was used to calculate the average value of the number of iterations, $kav$, which corresponds to the lower estimate of this average value.
+This means that the algorithm was halted because the maximum allowed number of $K_{max}$ iterations was reached.
+In this case, the value $K_{max} = 10^6$ was used to calculate the average value of the number of iterations, $k_{av}$, which corresponds to the lower estimate of this average value.
 The number of unsolved problems is indicated in parentheses.
 
-  As can be seen from Table \ref{table:average_iters}, the sequential GSA surpasses the DIRECT and DIRECTl methods in all classes of problems in terms of the average number of iterations.
+  As can be seen from Table \ref{table:average_iters}, the sequential GSA surpasses the DIRECT and DIRECT\textit{l} methods in all classes of problems in terms of the average number of iterations.
 At the same time, in the 5-Hard class, none of the methods solved all the problems: DIRECT failed to solve 16 problems, DIRECTl and GSA -- 4 problems each.
@@
 
-Следующий эксперимент будет проведен на 6-ти мерных задачах GKLS.
+%Следующий эксперимент будет проведен на 6-ти мерных задачах GKLS.
 В таблице \ref{table:GKLS6_average_MMAS_GPU} приведено ускорение по времени и по итерациям полученное при решении серии задач на GPU с использованием алгоритма объединяющего множественные развертки и адаптивную схему по сравнению с адаптивной схемой на CPU использующей 4 потока.
@@
 %
-\begin{thebibliography}{6}
+\bibliographystyle{spmpsci}
+\bibliography{bibliography}{}
 
-\bibitem {Ref1}	Hwu, W. GPU Computing Gems: Emerald Edition — San Francisco: Morgan Kaufmann, 2011. — 886 p.
-\bibitem {Ref2}	A.M.
+%\begin{thebibliography}{6}
+%
+%\bibitem {Ref1}	Hwu, W. GPU Computing Gems: Emerald Edition — San Francisco: Morgan Kaufmann, 2011. — 886 p.
+%\bibitem {Ref2}	A.M.
 Ferreiro, J.A.
@@
 863–890.
-\bibitem {Ref3}	J.M. Garcia-Martinez, E.M. Garzon, P.M. Ortigosa. A GPU implementation of a hybrid evolutionary algorithm: GPuEGO // Journal of super-computing. — 2014. — V. 70, No 2. — P. 684–695.
-\bibitem {Ref4}	Langdon, W.B. Graphics processing units and genetic programming: an overview // Soft Computing. — 2011. — V. 15, No 8. — P. 1657–1669.
-\bibitem {Ref5}	Ю.Г.
+%\bibitem {Ref3}	J.M. Garcia-Martinez, E.M. Garzon, P.M. Ortigosa. A GPU implementation of a hybrid evolutionary algorithm: GPuEGO // Journal of super-computing. — 2014. — V. 70, No 2. — P. 684–695.
+%\bibitem {Ref4}	Langdon, W.B. Graphics processing units and genetic programming: an overview // Soft Computing. — 2011. — V. 15, No 8. — P. 1657–1669.
+%\bibitem {Ref5}	Ю.Г.
 Евтушенко, В.У.
@@
 255–269.
-\bibitem {Ref6}	J.
+%\bibitem {Ref6}	J.
 He, A.
@@
 217–245.
-\bibitem {Ref7}	R.
+%\bibitem {Ref7}	R.
 Paulavicius, J.
@@
 487–498.
-\bibitem {Ref8}	V. Kindratenko (ed.). Numerical computations with GPUs — New York: Springer, 2014. — 405 p.
-\bibitem {Ref9}	Strongin, R.G. Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms / R.G. Strongin, Ya.D. Sergeyev – Kluwer Academic Publishers, 2000. – 704 p.
-\bibitem {Ref10}	Sergeyev, Ya.D. Introduction to global optimization exploiting space-filling curves / Ya.D. Sergeyev, R.G. Strongin, D. Lera – Springer, 2013. – 125 p.
-\bibitem {Ref11}	Стронгин, Р.Г.
+%\bibitem {Ref8}	V. Kindratenko (ed.). Numerical computations with GPUs — New York: Springer, 2014. — 405 p.
+%\bibitem {Ref9}	Strongin, R.G. Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms / R.G. Strongin, Ya.D. Sergeyev – Kluwer Academic Publishers, 2000. – 704 p.
+%\bibitem {Ref10}	Sergeyev, Ya.D. Introduction to global optimization exploiting space-filling curves / Ya.D. Sergeyev, R.G. Strongin, D. Lera – Springer, 2013. – 125 p.
+%\bibitem {Ref11}	Стронгин, Р.Г.
 Параллельные вычисления в задачах глобальной оптимизации / Р.Г.
@@
 – 280 с.
-\bibitem {Ref12}	Barkalov, K.A.
+%\bibitem {Ref12}	Barkalov, K.A.
 A global optimization technique with an adaptive order of checking for constraints / K.A.
@@
 1289-1300.
-\bibitem {Ref13}	Gergel, V.P.
+%\bibitem {Ref13}	Gergel, V.P.
 A method of using derivatives in the minimization of multiextremum functions / V.P.
@@
 729 – 742.
-\bibitem {Ref14}	Gergel, V.P.
+%\bibitem {Ref14}	Gergel, V.P.
 A global optimization algorithm for multivariate functions with lipschitzian first derivatives / V.P.
@@
 257-281.
-\bibitem {Ref15}	Gaviano, M.
+%\bibitem {Ref15}	Gaviano, M.
 Software for generation of classes of test functions with known local and global minima for global optimization/ M.
@@
 469-480.
-\bibitem {Ref16}	Сергеев, Я.Д. Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. – 352 c.
-\bibitem {Ref17}	Gablonsky, J.M. A Locally-Biased Form of the DIRECT Algorithm / J.M.  Gablonsky, C.T. Kelley, // Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.
-\bibitem {Ref18}	Sergeyev, Ya.D.
+%\bibitem {Ref16}	Сергеев, Я.Д. Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. – 352 c.
+%\bibitem {Ref17}	Gablonsky, J.M. A Locally-Biased Form of the DIRECT Algorithm / J.M.  Gablonsky, C.T. Kelley, // Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.
+%\bibitem {Ref18}	Sergeyev, Ya.D.
 Global search based on efficient diagonal partitions and a set of Lipschitz constants / Ya.D.
@@
 910–937.
-\bibitem {Ref19}	Strongin R.G. Algorithms for multi-extremal mathematical programming problems employ-ing the set of joint space-filling curves// J. of Global Optimization, 1992. No2. P. 357–378. 
-\bibitem {Ref20}	 Strongin R.G., Sergeyev Ya.D. Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht, 2000. 
-\bibitem {Ref21}	Barkalov K., Lebedev I., Sovrasov V.
+%\bibitem {Ref19}	Strongin R.G. Algorithms for multi-extremal mathematical programming problems employ-ing the set of joint space-filling curves// J. of Global Optimization, 1992. No2. P. 357–378. 
+%\bibitem {Ref20}	 Strongin R.G., Sergeyev Ya.D. Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht, 2000. 
+%\bibitem {Ref21}	Barkalov K., Lebedev I., Sovrasov V.
 Comparison of dimensionality reduction schemes for parallel global optimization algorithms.
 Communications in Computer and Information Science Volume 965, 2019, Pages 50-62
-\bibitem {Ref22}	Городецкий С.Ю., Гришагин В.А. Нелинейное программирование и многоэкстремальная оптимизация. Н.Новгород: Изд-во ННГУ, 2007.
-\bibitem {Ref23}	Grishagin V.A., Sergeyev Y.D., Strongin R.G. Parallel Characteristical Algorithms for Solving Problems of Global Optimization // J. Glob. Optim. 1997. V.~10~(2). P. 185--206.
-\bibitem {Ref24}	Sergeyev Y., Grishagin V. Parallel Asynchronous Global Search and the Nested Optimization Scheme // J. Comput. Anal. Appl. 2001. V. 3~(2). P. 123--145.
-\bibitem {Ref25}	Gergel V., Grishagin V., Israfilov R. Local Tuning in Nested Scheme of Global Optimization // Procedia Computer Sci. 2015. V.~51~(1). P.~865--874.
-\bibitem {Ref26}	Barkalov K., Gergel V.
+%\bibitem {Ref22}	Городецкий С.Ю., Гришагин В.А. Нелинейное программирование и многоэкстремальная оптимизация. Н.Новгород: Изд-во ННГУ, 2007.
+%\bibitem {Ref23}	Grishagin V.A., Sergeyev Y.D., Strongin R.G. Parallel Characteristical Algorithms for Solving Problems of Global Optimization // J. Glob. Optim. 1997. V.~10~(2). P. 185--206.
+%\bibitem {Ref24}	Sergeyev Y., Grishagin V. Parallel Asynchronous Global Search and the Nested Optimization Scheme // J. Comput. Anal. Appl. 2001. V. 3~(2). P. 123--145.
+%\bibitem {Ref25}	Gergel V., Grishagin V., Israfilov R. Local Tuning in Nested Scheme of Global Optimization // Procedia Computer Sci. 2015. V.~51~(1). P.~865--874.
+%\bibitem {Ref26}	Barkalov K., Gergel V.
 Multilevel Scheme of Dimensionality Reduction for Parallel Global Search Algorithms / OPT-i 2014 --- 1st Int.
@@
 2111--2124.
-\bibitem {Ref27}	Barkalov K., Lebedev I. Solving Multidimensional Global Optimization Problems using Graphics Accelerators // CCIS. 2016. V. 687. P. 224--235.
-\bibitem {Ref28}	Gergel V., Grishagin V., Gergel A. Adaptive Nested Optimization Scheme for Multidimensional Global Search // J. Glob. Optim. 2016. V.~66~(1). P.~35--51.
-\bibitem {Ref29}	Grishagin V., Israfilov R., Sergeyev Y. Comparative Efficiency of Dimensionality Reduction Schemes in Global Optimization // AIP Conf. Proc. 2016. V. 1776. Art. No. 060011.
-\bibitem {Ref30}	Grishagin V., Israfiov R., Sergeyev Y.
+%\bibitem {Ref27}	Barkalov K., Lebedev I. Solving Multidimensional Global Optimization Problems using Graphics Accelerators // CCIS. 2016. V. 687. P. 224--235.
+%\bibitem {Ref28}	Gergel V., Grishagin V., Gergel A. Adaptive Nested Optimization Scheme for Multidimensional Global Search // J. Glob. Optim. 2016. V.~66~(1). P.~35--51.
+%\bibitem {Ref29}	Grishagin V., Israfilov R., Sergeyev Y. Comparative Efficiency of Dimensionality Reduction Schemes in Global Optimization // AIP Conf. Proc. 2016. V. 1776. Art. No. 060011.
+%\bibitem {Ref30}	Grishagin V., Israfiov R., Sergeyev Y.
 Convergence Conditions and Numerical Comparison of Global Optimization Methods Based on Dimensionality Reduction Schemes // Appl.
@@
 270--280.
-\bibitem {Ref31}	Gergel V., Grishagin V., Israfilov R. Parallel Dimensionality Reduction for Multiextremal Optimization Problems // LNCS. 2019. V. 11657. P. 166--178.
-\bibitem {Ref32}	Стронгин Р.Г., Гергель В.П., Баркалов К.А. Адаптивная глобальная оптимизация на основе блочно-рекурсивной схемы редукции размерности.
-
-\end{thebibliography}
-
+%\bibitem {Ref31}	Gergel V., Grishagin V., Israfilov R. Parallel Dimensionality Reduction for Multiextremal Optimization Problems // LNCS. 2019. V. 11657. P. 166--178.
+%\bibitem {Ref32}	Стронгин Р.Г., Гергель В.П., Баркалов К.А. Адаптивная глобальная оптимизация на основе блочно-рекурсивной схемы редукции размерности. 
+%
+%
+%
+%\end{thebibliography}
+%
 \end{document}

### 2020-04-13 2020/RSD_GPU/RSD2020.tex
Обновленные таблицы. Поправлено оформление.

@@
 $\left\| {{y}^{k}}-{{y}^{*}} \right\|\le \delta $.
-Also, the size of the neighborhood was chosen (in accordance with \cite{Sergeyev2006}) as $\delta =\left\| b-a \right\|\sqrt[N]{\Delta }$, here $N$ is the dimension of the problem to be solved, a и b are the boundaries of the search domain $D$, the parameter  $\Delta $=10-6 for N=4 and $\Delta =10-7$ for $N=5$.
+Also, the size of the neighborhood was chosen (in accordance with \cite{Sergeyev2006}) as $\delta =\left\| b-a \right\|\sqrt[N]{\Delta }$, here $N$ is the dimension of the problem to be solved , a and b are the boundaries  of the search domain $D$, the parameter  $\Delta ={{10}^{-6}}$ for N=4 and $\Delta ={{10}^{-7}}$ for $N=5$.
 When using the AGP method for the Simple class, the parameter $r=4.5$ was selected, for the Hard class - $r=5.6$; the parameter for constructing the Peano curve was fixed, $m=10$.
@@
     \(N\) & Problem class & DIRECT & DIRECT\(l\) & GSA \\ \hline
-    4 & Simple & >47282 (4) & 18983 & 11953\\ \cline{2-5}
-      & Hard & >95708 (7) & 68754 & 25263\\ \hline
-    5 & Simple &>16057 (1) & 16758 & 15920\\ \cline{2-5}
-      & Hard & >217215 (16) & >269064 (4) & >148342 (4)\\ \hline
+    4 & Simple & \(>\)47282 (4) & 18983 & 11953\\ \cline{2-5}
+      & Hard & \(>\)95708 (7) & 68754 & 25263\\ \hline
+    5 & Simple &\(>\)16057 (1) & 16758 & 15920\\ \cline{2-5}
+      & Hard & \(>\)217215 (16) & \(>\)269064 (4) & \(>\)148342 (4)\\ \hline
     \end{tabular}
@@
    Table \ref{table:average_iters} shows the average number of iterations, $k_{av}$, that the method performed when solving a series of problems from these classes.
-The symbol ">" reflects a situation where not all problems of the class were solved by any method whatsoever.
+The symbol \(">"\) reflects a situation where not all problems of the class were solved by any method whatsoever.
 This means that the algorithm was halted because the maximum allowed number of $K_{max}$ iterations was reached.
@@
 \begin{table}
-    \centering
-    \begin{tabular}{|c|p{3cm}|c|c|c|}
-    \hline
-    \(N\) & Problem class & P=2 & P=4 & P=8 \\ \hline
-    4 & Simple  & 3,49 & 8,22  &  \\ \cline{2-5}
-      & Hard    & 2,35 & 7,65  &  \\ \hline
-    5 & Simple  & 1,87 & 6,59  &  \\ \cline{2-5}
-      & Hard    & 1,48 & 10,80  &   \\ \hline
+                \caption{Results on CPU}
+                \label{table:average_AS_CPU}
+                \center
+                \begin{tabular}{cccccccccccc}
+                               \hline\noalign{\smallskip}
+                & \multicolumn{5}{c}{ Acceleration of iterations} & & \multicolumn{5}{c}{ Time acceleration}\\
+                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
+                               $p$ & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
+                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
+                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
+                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
+                               $p=2$&   3,49 & 2,35 &   & 0,86 & 2,03 &   & 1,71 & 1,17 &   & 0,95 & 0,4 \\
+                               $p=4$&   8,22 & 7,65 &   & 6,74 & 8,56 &   & 2,01 & 1,9  &   & 1,85 & 1,53 \\
+                               \noalign{\smallskip}\hline
                 \end{tabular}
-    \caption{Acceleration of iterations on the CPU}
-    \label{table:average_iters_AS_CPU}
+\end{table}
+
+   Let us now evaluate the acceleration achieved using parallel GSA using an adaptive dimensionality reduction scheme based on the number $P$ of cores used.
+Table \ref{table:average_AS_CPU} shows the acceleration by iterations and the time acceleration obtained when solving a series of problems on the CPU, compared to GSA.
+
+\begin{table}
+                \caption{Results on CPU}
+                \label{table:average_MMAS_CPU}
+                \center
+                \begin{tabular}{cccccccccccc}
+                               \hline\noalign{\smallskip}
+                & \multicolumn{5}{c}{ Acceleration of iterations} & & \multicolumn{5}{c}{ Time acceleration}\\
+                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
+                               $p$ & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
+                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
+                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
+                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
+                               $p=2$&    6,62  & 3,23  &   & 2,13 & 6,07  &   & 3,22 & 1,48 &   & 0,66 & 1,93 \\
+                               $p=4$&    21,46 &  9,97 &   & 6,9  & 16,27 &   & 5,23 & 2,42 &   & 0,85 & 1,9 \\
+                               \noalign{\smallskip}\hline
+                \end{tabular}
+\end{table}
+
+   Table \ref{table:average_MMAS_CPU} shows the acceleration in iterations and  the acceleration in time of the algorithm that combines multiple sweeps and an adaptive scheme for solving a series of problems on the CPU, compared to the sequential launch of the GSA.
+Two scans and, accordingly, two processes were used; each process used $P$ threads, and calculations were performed on a single cluster node.
+
+\begin{table}
+                \caption{Results on GPU}
+                \label{table:average_AS_GPU}
+                \center
+                \begin{tabular}{cccccccccccc}
+                               \hline\noalign{\smallskip}
+                & \multicolumn{5}{c}{ Acceleration of iterations} & & \multicolumn{5}{c}{ Time acceleration}\\
+                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
+                               $p$ & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
+                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
+                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
+                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
+                               $p=64$  &    5,06  & 3,93  &   & 1,19 	 & 5,41  &   & 2,2	    & 1,92	 &   & 0,47	& 2,21 \\
+                               $p=256$ &    19,91 & 14,95 &   & 11,9 	 & 23,36 &   & 8,28	& 6,91	     &   & 3,4	    & 6,1 \\
+							   $p=1024$&    15,62 & 52,91 &   & 22,12	 & 55,22 &   & 5,21	& 19,95      &   & 2,18	& 4,82 \\
+                               \noalign{\smallskip}\hline
+                \end{tabular}
 \end{table}
@@
 \begin{table}
-    \centering
-    \begin{tabular}{|c|p{3cm}|c|c|c|}
-    \hline
-    \(N\) & Problem class & P=2 & P=4 & P=8 \\ \hline
-    4 & Simple  & 1,71 & 2,01  &  \\ \cline{2-5}
-      & Hard    & 1,17 & 1,90  &  \\ \hline
-    5 & Simple  & 0,88 & 1,48  &  \\ \cline{2-5}
-      & Hard    & 0,70 & 2,39  &   \\ \hline
+                \caption{Results on GPU}
+                \label{table:average_MMAS_GPU}
+                \center
+                \begin{tabular}{cccccccccccc}
+                               \hline\noalign{\smallskip}
+                & \multicolumn{5}{c}{ Acceleration of iterations} & & \multicolumn{5}{c}{ Time acceleration}\\
+                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
+                               $p$ & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
+                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
+                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
+                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
+                               $p=64$  &    11,32	 & 6,83	 &   & 8,19	     & 10,66  &   & 2,44	 & 1,63	 &   & 1,33	 & 1,58 \\
+                               $p=256$ &    38,96	 & 31,52 &   & 25,43	 & 48,23  &   & 8,01	 & 7,04	 &   & 2,31	 & 3,42 \\
+							   $p=1024$&    128,39	 & 83,46 &   & 98,43	 & 146,23 &   & 20,45	 & 11,06 &   & 2,75	 & 2,48 \\
+                               \noalign{\smallskip}\hline
                 \end{tabular}
-    \caption{Time acceleration on the CPU}
-    \label{table:average_Time_AS_CPU}
 \end{table}
 
-   Let us now evaluate the acceleration achieved using parallel GSA using an adaptive dimensionality reduction scheme based on the number $P$ of cores used.
-Table \ref{table:average_iters_AS_CPU} shows the acceleration by iterations, and Table \ref{table:average_Time_AS_CPU} shows the time acceleration obtained when solving a series of problems on the CPU, compared to GSA.
-
-\begin{table}
-    \centering
-    \begin{tabular}{|c|p{3cm}|c|c|c|}
-    \hline
-    \(N\) & Problem class & P=2 & P=4 & P=8 \\ \hline
-    4 & Simple &   6,62 & 21,46  &  \\ \cline{2-5}
-      & Hard   &   3,23 & 9,97  &  \\ \hline
-    5 & Simple &   3,37 & 11,20  &  \\ \cline{2-5}
-      & Hard   &   2,52 & 8,54  &   \\ \hline
-    \end{tabular}
-    \caption{Time acceleration on the CPU}
-    \label{table:average_iters_MMAS_CPU}
-\end{table}
-
-\begin{table}
-    \centering
-    \begin{tabular}{|c|p{3cm}|c|c|c|}
-    \hline
-    \(N\) & Problem class & P=2 & P=4 & P=8 \\ \hline
-    4 & Simple &   3,22 & 5,23  &  \\ \cline{2-5}
-      & Hard   &   1,48 & 2,42  &  \\ \hline
-    5 & Simple &   1,30 & 1,97  &  \\ \cline{2-5}
-      & Hard   &   0,68 & 0,88  &   \\ \hline
-    \end{tabular}
-    \caption{Acceleration of iterations on the CPU}
-    \label{table:average_Time_MMAS_CPU}
-\end{table}
-
-   Table \ref{table:average_iters_MMAS_CPU} shows the acceleration in iterations, and Table \ref{table:average_Time_MMAS_CPU} shows the acceleration in time of the algorithm that combines multiple sweeps and an adaptive scheme for solving a series of problems on the CPU, compared to the sequential launch of the GSA.
-Two scans and, accordingly, two processes were used; each process used $P$ threads, and calculations were performed on a single cluster node.
-
-\begin{table}
-    \centering
-    \begin{tabular}{|c|p{3cm}|c|c|c|}
-    \hline
-    \(N\) & Problem class & P=64 & P=256 & P=1024 \\ \hline
-    4 & Simple &   5,06 & 19,91 & 15,62  \\ \cline{2-5}
-      & Hard   &   3,93 & 14,95 &   \\ \hline
-    5 & Simple &   4,16 & 16,12 & 67,22  \\ \cline{2-5}
-      & Hard   &   3,36 & 10,02 & 42,04   \\ \hline
-    \end{tabular}
-    \caption{Acceleration of iterations on the GPU}
-    \label{table:average_iters_AS_GPU}
-\end{table}
-
-\begin{table}
-    \centering
-    \begin{tabular}{|c|p{3cm}|c|c|c|}
-    \hline
-    \(N\) & Problem class & P=64 & P=256 & P=1024 \\ \hline
-    4 & Simple &   2,20 & 8,28 & 5,21  \\ \cline{2-5}
-      & Hard   &   1,92 & 6,91 &   \\ \hline
-    5 & Simple &   1,58 & 4,18 & 8,23  \\ \cline{2-5}
-      & Hard   &   1,22 & 2,39 & 3,74   \\ \hline
-    \end{tabular}
-    \caption{Time acceleration on the GPU}
-    \label{table:average_Time_AS_GPU}
-\end{table}
-
-\begin{table}
-    \centering
-    \begin{tabular}{|c|p{3cm}|c|c|c|}
-    \hline
-    \(N\) & Problem class & P=64 & P=256 & P=1024 \\ \hline
-    4 & Simple &   11,32 & 38,96 & 128,39  \\ \cline{2-5}
-      & Hard   &   6,83  & 31,52 &   \\ \hline
-    5 & Simple &   4,99  & 39,37 & 80,56  \\ \cline{2-5}
-      & Hard   &   4,61  & 17,28 & 113,61   \\ \hline
-    \end{tabular}
-    \caption{Acceleration of iterations on the GPU}
-    \label{table:average_iters_MMAS_GPU}
-\end{table}
-
-\begin{table}
-    \centering
-    \begin{tabular}{|c|p{3cm}|c|c|c|}
-    \hline
-    \(N\) & Problem class & P=64 & P=256 & P=1024 \\ \hline
-    4 & Simple &   2,44 & 8,01 & 20,45  \\ \cline{2-5}
-      & Hard   &   1,63 & 7,04 &   \\ \hline
-    5 & Simple &   0,77 & 3,97 & 2,37  \\ \cline{2-5}
-      & Hard   &   0,61 & 0,96 & 2,11   \\ \hline
-    \end{tabular}
-    \caption{Time acceleration on the GPU}
-    \label{table:average_Time_MMAS_GPU}
-\end{table}
-
-Tables \ref{table:average_iters_AS_GPU} and  \ref{table:average_Time_AS_GPU} show the acceleration obtained when solving a series of problems on the GPU using an adaptive scheme compared to a similar launch on the CPU using 4 threads.
-Tables \ref{table:average_iters_MMAS_GPU} and \ref{table:average_Time_MMAS_GPU} show the acceleration of the algorithm combining multiple sweeps and an adaptive circuit when solving a series of problems on the GPU, compared to an adaptive scheme on the CPU using 4 threads.
-Two sweeps and, accordingly, two processes were used; each process used P threads, and calculations were performed on a single cluster node.
+Tables \ref{table:average_AS_GPU} show the acceleration obtained when solving a series of problems on the GPU using an adaptive scheme compared to a similar launch on the CPU using 4 threads.
+Tables \ref{table:average_MMAS_GPU} show the acceleration of the algorithm combining multiple sweeps and an adaptive circuit when solving a series of problems on the GPU, compared to an adaptive scheme on the CPU using 4 threads.
+Two sweeps and, accordingly, two processes were used; each process used $P$ threads, and calculations were performed on a single cluster node.
 

### 2020-04-14 2020/RSD_GPU/RSD2020.tex
fix one table
fix a lot of misprints

@@
 \usepackage{amssymb}
+\usepackage{cite}
 
@@
 
-Thus, calculating the optimization criterion can be implemented on the GPU, and the role of the optimization algorithm (running on the CPU) consists in the effective selection of points for conducting parallel tests.
-This scheme of working with the accelerator is fully consistent with the work of the parallel global search algorithm developed at the Lobachevsky State University of Nizhni Novgorod and presented in a series of publications [references].
+Thus, calculating the optimization criterion can be implemented on the GPU, and the role of the optimization algorithm (running on the CPU) consists in the effective selection of points for conducting parallel trials.
+This scheme of working with the accelerator is fully consistent with the work of the parallel global search algorithm developed at the Lobachevsky State University of Nizhni Novgorod and presented in a series of publications \cite{Barkalov2014,Barkalov2020,Barkalov2019,Grishagin2016,Grishagin2015,Grishagin2018,Strongin2000}.
 
 \section{Multidimensional parallel global search algorithm}
-Consider the problem of finding the global minimum of the \(N\)-dimensional function \(\varphi(y)\) in the hyperinterval  \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\)  We will assume that the function satisfies the Lipschitz condition with an a priori unknown constant \(L\).
+Consider the problem of finding the global minimum of the \(N\)-dimensional function \(\varphi(y)\) in the hyperinterval  \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
+We will assume that the function satisfies the Lipschitz condition with an a priori unknown constant \(L\).
 \begin{equation}
 \label{task}
-\varphi(y^*)=\min\{\varphi(y):y\in D\}
+\varphi(y^*)=\min\{\varphi(y):y\in D\},
 \end{equation}
@@
 \label{lip}
-|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
+|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty.
 \end{equation}
@@
 \label{cube}
-\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
+\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}.
 \end{equation}
 
-The questions of numerical construction of approximations to Piano curve (evolvents) and the corresponding theory are discussed in detail in \cite{Strongin2000}, \cite{Sergeyev2013}.
+The questions of numerical construction of approximations to Piano curve (evolvents) and the corresponding theory are discussed in detail in \cite{Sergeyev2013,Strongin2000}.
 Using evolvents \(y(x)\) reduces the multidimensional problem (\ref{task}) to a one-dimensional problem
@@
 \label{oneDimTask}
-\varphi(y^*)=\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}
+\varphi(y^*)=\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
 \end{displaymath}
 
-An important property is that the relative differences of the function remain limited: if the function \(\varphi(y)\) in the region \(D
-\) satisfies the Lipschitz condition, then the function \(\varphi(y(x))\) in the interval \([0,1]\) will satisfy the uniform Holder condition
+An important property is that the relative differences of the function remain limited: if the function \(\varphi(y)\) in the region \(D\) satisfies the Lipschitz condition, then the function \(\varphi(y(x))\) in the interval \([0,1]\) will satisfy a uniform H{\"o}lder condition
 \begin{displaymath}
 \label{holder}
-|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
+|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
 \end{displaymath}
-where the Holder constant \(H\) is related to the Lipschitz constant \(L\) by the ratio
-\begin{displaymath}
-H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}
-\end{displaymath}
-
-Therefore, without limiting generality, we can consider minimizing the one-dimensional function \(f(x)=\varphi(y(x)), x\in[0,1]\), which satisfies the Holder condition.
+where the H{\"o}lder constant \(H\) is related to the Lipschitz constant \(L\) by the ratio $ H=2L\sqrt{N+3}$. 
+Therefore, without limiting generality, we can consider minimizing the one-dimensional function \(f(x)=\varphi(y(x)), x\in[0,1]\), which satisfies the H{\"o}lder condition.
 
@@
 We will call the process of computing the value of a function at a single point a trial.
-Assume that we have \(P\geqslant 1\) computational elements at our disposal and \(P\) tests which are performed simultaneously (synchronously) within a single iteration of the method.
-Let \(k(n)\) denote the total number of tests performed, after \(n\) parallel iterations.
+Assume that we have \(p\geqslant 1\) computational elements at our disposal and \(p\) trials which are performed simultaneously (synchronously) within a single iteration of the method.
+Let \(k(n)\) denote the total number of trials performed, after \(n\) parallel iterations.
 
-In the first iteration of the method, the test is carried out at an arbitrary internal point \(x^1\) of the interval \([0,1]\).
-Let \(n>1\) iterations of the method be performed, during which tests were carried out at \(k = k(n)\) points \(x^i, 1\leqslant i\leqslant k\).
-Then the points  \(x^{k+1},\dotsc,x^{k+p}\).
-search tests of the following \((n+1)\)-th iteration are determined in accordance with the rules:
+In the first iteration of the method, the trial is carried out at an arbitrary internal point \(x^1\) of the interval \([0,1]\).
+Let \(n>1\) iterations of the method be performed, during which trials were carried out at \(k = k(n)\) points \(x^i, 1\leqslant i\leqslant k\).
+Then the trial points  \(x^{k+1},\dotsc,x^{k+p}\) of the next \((n+1)\)-th iteration are determined in accordance with the following rules.
 
 Step 1.
-Renumber the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which includes the boundary points of the interval \([0,1]\), as well as the points of the previous tests, with the lower indices in the order of their increasing coordinate values, i.e.
+Renumber the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which includes the boundary points of the interval \([0,1]\), as well as the points of the previous trials, with the lower indices in the order of their increasing coordinate values, i.e.
 \begin{displaymath}
-0=x_0<x_1<\dotsc<x_{k+1}=1
+0=x_0<x_1<\dotsc<x_{k+1}=1.
 \end{displaymath}
@@
 \label{step3_1}
-R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M}
+R(1)=2\Delta_1-4\dfrac{z_1}{M}, \; R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
 \end{displaymath}
@@
 \label{step3_2}
-R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1
+R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1.
 \end{displaymath}
@@
 \end{displaymath}
-and select \(P\) of the largest characteristics with interval numbers \(t_j,1\leqslant j\leqslant P\).
+and select \(p\) of the largest characteristics with interval numbers \(t_j,1\leqslant j\leqslant p\).
 
-Step 5. Carry out new tests at the points \(x_{k+j},1\leqslant j\leqslant P\), calculated using the formulas
+Step 5. Carry out new trials at the points \(x_{k+j},1\leqslant j\leqslant p\), calculated using the formulas
 \begin{displaymath}
-x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},t_j=1,t_j=k+1
+x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},\; t_j=1, \; t_j=k+1
 \end{displaymath}
@@
 \label{step5}
-{{x}^{k+1}}=\frac{{{x}_{{{t}_{j}}}}+{{x}_{{{t}_{j}}-1}}}{2}-\text{sign}({{z}_{{{t}_{j}}}}-{{z}_{{{t}_{j}}-1}})\frac{1}{2r}{{\left[ \frac{|{{z}_{{{t}_{j}}}}-{{z}_{{{t}_{j}}-1}}|}{\mu } \right]}^{N}}, 1<{{t}_{j}}<k+1
+{{x}^{k+1}}=\frac{{{x}_{{{t}_{j}}}}+{{x}_{{{t}_{j}}-1}}}{2}-\text{sign}({{z}_{{{t}_{j}}}}-{{z}_{{{t}_{j}}-1}})\frac{1}{2r}{{\left[ \frac{|{{z}_{{{t}_{j}}}}-{{z}_{{{t}_{j}}-1}}|}{\mu } \right]}^{N}}, 1<{{t}_{j}}<k+1.
 \end{displaymath}
 
-The algorithm stops working if the condition \(\Delta_{t_j}\leqslant \varepsilon\)  is satisfied for at least one number \(t_j,1\leqslant j\leqslant P\); here \(\varepsilon>0\)  is the specified accuracy.
+The algorithm stops working if the condition \(\Delta_{t_j}\leqslant \varepsilon\)  is satisfied for at least one number \(t_j,1\leqslant j\leqslant p\); here \(\varepsilon>0\)  is the specified accuracy.
 As an estimate of the globally optimal solution to the problem (\ref{task}), the values are selected
 \begin{displaymath}
-f_{k}^{*}=\underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}}),   x_{k}^{*}=\arg \quad \underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}})
+f_{k}^{*}=\underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}}),   x_{k}^{*}=\arg \quad \underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}}).
 \end{displaymath}
 
-For the rationale in using this method of organizing parallel computing, see \cite{Ref11}.
-Modifications that account for the presence of inequality constraints in the problem, as well as information about the derivative of the objective function, are presented in \cite{Ref12} – \cite{Ref14}.
+For the rationale in using this method of organizing parallel computing see \cite{Strongin2000}. 
+%Modifications that account for the presence of inequality constraints in the problem, as well as information about the derivative of the objective function, are presented in \cite{Ref12} – \cite{Ref14}.
 
@@
 A possible way to overcome this disadvantage is to use a set of evolvents (multiple mappings)
-
 \begin{displaymath}
@@
 \end{displaymath}
-
-instead of using a single Peano curve \(y(x)\) (see \cite{Strongin1992}, \cite{Strongin2000}).
+instead of using a single Peano curve \(y(x)\) (see \cite{Strongin1992,Strongin2000}).
 For example, each Peano curve \(y^i(x)\) from \(Y_L(x)\) can be obtained as a result of some shift along the main diagonal of the hyperinterval \(D\), the original curve \(y(x)\).
-Another way is to rotate the sweep around the coordinate.
-The set of evolvents that have been constructed allows us to obtain for any close images \(y', y''\), close prototypes \(x', x''\) for some mapping \(y^i(x)\).
+Another way is to rotate the evolvent around the origin.
+The set of evolvents that have been constructed allows us to obtain for any close images \(y', y''\), close preimages \(x', x''\) for some mapping \(y^i(x)\).
 
 Using a set of mappings leads to the formation of a corresponding set of one-dimensional multiextremal problems
-
 \begin{displaymath}
 \label{6_problem_l}
-\min{\left\{\varphi(y^l(x)):x\in [0,1], \;  \right\}}, \ 0 \leq l \leq L.
+\min{\left\{\varphi(y^l(x)):x\in [0,1], \;  \right\}}, \ 0 \leqslant l \leqslant L.
 \end{displaymath}
-
 Each problem from this set can be solved independently, and any calculated value \(z= \varphi(y')\), \(y'=y^i(x')\) of the function \(\varphi(y)\) in the \(i\)-th problem can be interpreted as calculating the value \(z= \varphi(y')\), \(y'=y^s(x'')\) for any other \(s\)-th problem without repeated labor-intensive calculations of the function \(\varphi(y)\).
@@
 
-The recursive optimization scheme is based on the well-known relation \cite{Ref22}
+The recursive optimization scheme is based on the well-known relation
 \begin{equation}
 \label{nestedScheme}
-\min{\varphi(y):y\in D}=\min_{a_1\leqslant y_1\leqslant b_1}\min_{a_2\leqslant y_2\leqslant b_2}\dots\min_{a_1\leqslant y_N\leqslant b_N}\varphi(y)
+\min{\varphi(y):y\in D}=\min_{a_1\leqslant y_1\leqslant b_1}\min_{a_2\leqslant y_2\leqslant b_2}\dots\min_{a_1\leqslant y_N\leqslant b_N}\varphi(y),
 \end{equation}
@@
 \label{phiN}
-\varphi_N(y_1,\dots,y_N)=\varphi(y_1,\dots,y_N)
+\varphi_N(y_1,\dots,y_N)=\varphi(y_1,\dots,y_N),
 \end{equation}
@@
 \label{phiI_}
-\varphi_i(y_1,\dots,y_i)=\min_{a_{i+1}\leqslant y_{i+1} \leqslant b_{i+1}}\varphi_{i+1}(y_1,\dots,y_i,y_{i+1}),1\leqslant i\leqslant N-1
+\varphi_i(y_1,\dots,y_i)=\min_{a_{i+1}\leqslant y_{i+1} \leqslant b_{i+1}}\varphi_{i+1}(y_1,\dots,y_i,y_{i+1}),1\leqslant i\leqslant N-1.
 \end{equation}
 
-Then, in accordance with the relation (\ref{nestedScheme}) , the solution of the original problem (\ref{task}) is reduced to the solution of a one-dimensional problem
-
+Then, in accordance with the relation (\ref{nestedScheme}), the solution of the original problem (\ref{task}) is reduced to the solution of a one-dimensional problem
 \begin{equation}
 \label{phiFirst}
-\varphi_1(y_1^*)=\min\{\varphi_1(y_1),y_1\in [a,b]\}
+\varphi_1(y_1^*)=\min\{\varphi_1(y_1),y_1\in [a,b]\}.
 \end{equation}
@@
 However, each calculation of the value of the one-dimensional function  \(\varphi_1(y_1)\) at some fixed point corresponds to the solution of a one-dimensional minimization problem 
-
 \begin{displaymath}
 \label{phi2}
-\varphi_2(y_1,y_2^*)=\min\{\varphi(y_1,y_2):y_2\in [a_2,b_2]\}
+\varphi_2(y_1,y_2^*)=\min\{\varphi(y_1,y_2):y_2\in [a_2,b_2]\}.
 \end{displaymath}
-
 And so on, until the calculation   according to (\ref{phiN}).
 
-For the recursive scheme described above, a generalization (block recursive scheme) is proposed that combines the use of evolvents and a recursive scheme in order to efficiently parallelize computations.
+For the recursive scheme described above, a generalization (\textit{block recursive scheme}) is proposed that combines the use of evolvents and a recursive scheme in order to efficiently parallelize computations.
 
@@
 Using new variables, the main relation of the multistep scheme (\ref{nestedScheme}) can be rewritten as
-
 \begin{equation}
 \label{blockNested}
-\min_{y\in D} \varphi(y)=\min_{u_1\in D_1}\min_{u_2\in D_2}\dots\min_{u_M\in D_M}\varphi(y)
+\min_{y\in D} \varphi(y)=\min_{u_1\in D_1}\min_{u_2\in D_2}\dots\min_{u_M\in D_M}\varphi(y),
 \end{equation}
+where the subdomains  \(D_i,1\leqslant i\leqslant M\), are projections of the original search domain \(D\) onto the subspaces corresponding to the variables \(u_1,1\leqslant i\leqslant M\).
 
-where the subdomains  \(D_i,1\leqslant i\leqslant M\), are projections of the original search domain \(D\) onto the subspaces corresponding to the variables \(u_1,1\leqslant i\leqslant M\) .
-
-The formulas that determine the method for solving problem (\ref{task}) based on relations (\ref{blockNested}) generally coincide with the recursive scheme (\ref{phiN})-(\ref{phiFirst}).
+The formulas that determine the method for solving problem (\ref{task}) based on relations (\ref{blockNested}) generally coincide with the recursive scheme (\ref{phiN})--(\ref{phiFirst}).
 One need only replace the original variables  \(y_i,1\leqslant i\leqslant N\), with block variables  \(u_1,1\leqslant i\leqslant M\).
-In this case, the fundamental difference from the original scheme is the fact that the block diagram has nested subproblems
-
+In this case, the fundamental difference from the original scheme is the fact that the block scheme has nested subproblems
 \begin{equation}
 \label{subTasks}
-\varphi_i(u_1,\dots,u_i)=\min_{u_{i+1}\in D_{i+1}}\varphi_{i+1}(u_1,\dots,u_i,u_{i+1}),1\leqslant i\leqslant M-1
+\varphi_i(u_1,\dots,u_i)=\min_{u_{i+1}\in D_{i+1}}\varphi_{i+1}(u_1,\dots,u_i,u_{i+1}),1\leqslant i\leqslant M-1,
 \end{equation}
-
 which are multidimensional, and to solve them, a method of reducing dimensionality based on Peano curves can be applied.
@@
 The solution to the resulting set of subproblems (\ref{subTasks}) can be organized in various ways.
-
-The obvious method (elaborated in detail in \cite{Grishagin1997,Grishagin2001,Grishagin2015} for the nested optimization scheme and in \cite{Barkalov2014},\cite{Ref27} for the block nested optimization scheme) is based on solving subproblems in accordance with the recursive order of their generation.
+The obvious method (elaborated in detail in \cite{Grishagin2015} for the nested optimization scheme and in \cite{Barkalov2014} for the block nested optimization scheme) is based on solving subproblems in accordance with the recursive order of their generation.
 However, there is a significant loss of information about the target function.
@@
 Another approach is an adaptive scheme in which all subproblems are solved simultaneously, which makes it possible to more fully take into account information about a multidimensional problem and thereby to speed up the process of solving it.
-
-In the case of one-dimensional subproblems, this approach was theoretically substantiated and tested in \cite{Grishagin2016,Grishagin2016_1,Grishagin2018}, and in the article [] a generalization of the adaptive scheme for multidimensional subproblems was proposed.
+In the case of one-dimensional subproblems, this approach was theoretically substantiated and tested in \cite{Grishagin2016,Grishagin2018}, and in the paper \cite{Barkalov2020} a generalization of the adaptive scheme for multidimensional subproblems was proposed.
 
@@
 Under the new scheme:
-
-•	to calculate the value of the function of the \(i\)-th level from (\ref{subTasks}), a new level \(i+1\) problem is generated in which tests are carried out, after which a new generated problem is included in the set of existing problems to be solved;
-
-•	the iteration of the global search consists in choosing \(P\) (the most promising) problem from the set of existing problems in which tests are carried out; points for new tests are determined in accordance with the algorithm of subsection 1;
-
-•	the minimum values of functions from (\ref{subTasks}) are their current estimates based on the accumulated search information. .
+	\begin{itemize}
+		\item to calculate the value of the function of the \(i\)-th level from (\ref{subTasks}), a new \(i+1\)-th level problem is generated in which trials are carried out, after which a new generated problem is included in the set of existing problems to be solved;
+		\item the iteration of the global search consists in choosing \(p\) (the most promising) problem from the set of existing problems in which trials are carried out; points for new trials are determined in accordance with the algorithm of subsection 1;
+		\item the minimum values of functions from (\ref{subTasks}) are their current estimates based on the accumulated search information.
+	\end{itemize}
 
 A brief description of the main steps of a block adaptive dimensionality reduction scheme is as follows.
+Let the nested subproblems in the form (\ref{subTasks}) be solved using the global search algorithm described in Section 2.
+Then each subproblem (\ref{subTasks}) can be assigned a numerical value called the characteristic of this problem.
+As such, we can take the maximum characteristic \(R(t)\) of the intervals formed in this problem.
+In accordance with the rule for calculating characteristics, the higher the value of the characteristic, the more promising the subproblem is in the continued search for the global minimum of the original problem (\ref{task}).
 
-Let the nested subproblems in the form (\ref{subTasks}) be solved using the global search algorithm described in subsection 1.
-Then each subproblem (\ref{subTasks}) can be assigned a numerical value called the characteristic of this problem.
-As such, we can take the maximum characteristic of the intervals formed in this problem – \(R(t)\).
-In accordance with the rule for calculating characteristics, the higher the value of the characteristic, the more promising the subproblem is in the continued search for the global minimum of the original problem (\ref{task}).
-Therefore, at each iteration, subproblems with the maximum characteristic are selected for conducting the next test.
-The test either leads to the calculation of the value of the objective function \(\varphi_1(y)\) (if the selected subproblem belonged to the level \(j=M\)), or generates new subproblems according to (\ref{subTasks}) for \(j \geqslant  M-1\).
+Therefore, at each iteration, subproblems with the maximum characteristic are selected for conducting the next trial.
+The trial either leads to the calculation of the value of the objective function \(\varphi_1(y)\) (if the selected subproblem belonged to the level \(j=M\)), or generates new subproblems according to (\ref{subTasks}) for \(j \geqslant  M-1\).
 In the latter case, the newly generated problems are added to the current set of problems, their characteristics are calculated, and the process is repeated.
 The optimization process is completed when the root problem satisfies the condition for stopping the algorithm that solves this problem.
-Some results pointing in this direction are presented in \cite{Ref31}, \cite{Ref32}.
+Some results pointing in this direction are presented in \cite{Barkalov2020}.
 
@@
 Naturally, this requires implementing a procedure for calculating the value of a function on the GPU.
-Data transfers from the CPU to the GPU will be minimal: one need only transfer the coordinates of the test points to the GPU, and get back the function values at these points.
-Functions that determine the processing of test results in accordance with the algorithm, and require work with large amounts of accumulated search information, can be effectively implemented on the CPU.
+Data transfers from the CPU to the GPU will be minimal: one need only transfer the coordinates of the trial points to the GPU, and get back the function values at these points.
+Functions that determine the processing of trial results in accordance with the algorithm, and require work with large amounts of accumulated search information, can be effectively implemented on the CPU.
 
    The general scheme for organizing calculations using the GPU is shown in \ref{fig:GPU_Algorithm}.
-In accordance with this scheme, steps 1-4 of the parallel global search algorithm are performed on the CPU.
-The coordinates of the P test points calculated in step 4 of the algorithm are accumulated in the intermediate buffer and then transmitted to the GPU.
-On the GPU, the function values are calculated at these points, after which the test results (again through the intermediate buffer) are transferred to the CPU.
+In accordance with this scheme, steps 1--4 of the parallel global search algorithm are performed on the CPU.
+The coordinates of the $p$ trial points calculated in step 4 of the algorithm are accumulated in the intermediate buffer and then transmitted to the GPU.
+On the GPU, the function values are calculated at these points, after which the trial results (again through the intermediate buffer) are transferred to the CPU.
 
@@
 This subproblem will be reduced to a set of one-dimensional problems that will be solved in parallel, each in a separate process.
-Test results at point x obtained for the problem solved by a specific processor are interpreted as the test results in the remaining problems (at the corresponding points \(u^1_1,… , u^s_1\) )
+Trial results at point $x$ obtained for the problem solved by a specific processor are interpreted as the trial results in the remaining problems (at the corresponding points \(u^1_1,..., u^s_1\) ).
 Then, applying an adaptive scheme to solve the nested subproblems (\ref{subTasks}), we get a parallel algorithm with a wide degree of variability.
@@
 
-The subproblem of the last \((M - 1)\) level \({{\varphi }_{M-1}}({{u}_{1}},...,{{u}_{M-1}})=\underset{{{u}_{M}}\in {{D}_{M}}}{\mathop{\min }}\,{{\varphi }_{M}}({{u}_{1}},...,{{u}_{M}})\) differs from all the previous subproblems – it calculates the values of the optimized function, since ${{\varphi }_{M}}({{u}_{1}},...,{{u}_{M}})=\varphi ({{y}_{1}},...,{{y}_{N}})$.
+The subproblem of the last \((M - 1)\)-th level 
+\[
+{{\varphi }_{M-1}}({{u}_{1}},...,{{u}_{M-1}})=\underset{{{u}_{M}}\in {{D}_{M}}}{\mathop{\min }}\,{{\varphi }_{M}}({{u}_{1}},...,{{u}_{M}})
+\]
+ differs from all the previous subproblems; it calculates the values of the objective function, since ${{\varphi }_{M}}({{u}_{1}},...,{{u}_{M}})=\varphi ({{y}_{1}},...,{{y}_{N}})$.
 This subproblem transfers between the CPU and the GPU.
@@
 
-In the study \cite{Gaviano2003}, \cite{Ref16}, a GKLS generator is described that allows one to generate multiextremal optimization problems with previously known properties: the number of local minima, the size of their regions of attraction, the global minimum point, the value of the function in it, etc.
+In the paper \cite{Gaviano2003} a GKLS generator is described that allows one to generate multiextremal optimization problems with previously known properties: the number of local minima, the size of their regions of attraction, the global minimum point, the value of the function in it, etc.
 
-Below are the results of a numerical comparison of three sequential algorithms -- DIRECT \cite{Jones2009}, DIRECT\textit{l} \cite{Gablonsky2001} and Global Search Algorithm (GSA); (results of the first two algorithms are given in the article \cite{Sergeyev2006}).
-A numerical comparison was carried out on the Simple and Hard function classes of dimension 4 and 5 from \cite{Sergeyev2006}.
-The Global minimum $y*$ was considered found if the algorithm generated a test point $y^k$ in the $\delta$-neighborhood of the global minimum, i.e.
-$\left\| {{y}^{k}}-{{y}^{*}} \right\|\le \delta $.
+Below are the results of a numerical comparison of three sequential algorithms -- DIRECT \cite{Jones2009}, DIRECT\textit{l} \cite{Gablonsky2001} and Global Search Algorithm (GSA) from Section 2.
+A numerical comparison was carried out on the Simple and Hard function classes of dimension 4 and 5 from \cite{Gaviano2003}.
+The global minimum $y*$ was considered found if the algorithm generated a trial point $y^k$ in the $\delta$-neighborhood of the global minimum, i.e.
+$\left\| {{y}^{k}}-{{y}^{*}} \right\| \leqslant \delta $.
 Also, the size of the neighborhood was chosen (in accordance with \cite{Sergeyev2006}) as $\delta =\left\| b-a \right\|\sqrt[N]{\Delta }$, here $N$ is the dimension of the problem to be solved , a and b are the boundaries  of the search domain $D$, the parameter  $\Delta ={{10}^{-6}}$ for N=4 and $\Delta ={{10}^{-7}}$ for $N=5$.
-When using the AGP method for the Simple class, the parameter $r=4.5$ was selected, for the Hard class - $r=5.6$; the parameter for constructing the Peano curve was fixed, $m=10$.
+When using the GSA method for the Simple class, the parameter $r=4.5$ was selected, for the Hard class $r=5.6$; the parameter for constructing the Peano curve was fixed, $m=10$.
 The maximum number of iterations allowed was $K_{max} = 10^6$.
@@
 \begin{table}
-    \centering
-    \begin{tabular}{|c|p{3cm}|c|c|c|}
-    \hline
-    \(N\) & Problem class & DIRECT & DIRECT\(l\) & GSA \\ \hline
-    4 & Simple & \(>\)47282 (4) & 18983 & 11953\\ \cline{2-5}
-      & Hard & \(>\)95708 (7) & 68754 & 25263\\ \hline
-    5 & Simple &\(>\)16057 (1) & 16758 & 15920\\ \cline{2-5}
-      & Hard & \(>\)217215 (16) & \(>\)269064 (4) & \(>\)148342 (4)\\ \hline
+	\caption{Average number of iterations $k_{av}$}\label{table:average_iters}
+	\center
+	\begin{tabular}{ccccc}
+		\hline\noalign{\smallskip}
+  	 $N$ & Problem class & DIRECT & DIRECT\textit{l} & GSA \\
+		\noalign{\smallskip} \hline \noalign{\smallskip}
+			4 &	Simple& $>$47282(4) &	18983 &	11953 \\
+	      & Hard &	$>$95708(7) &	68754 &	25263 \\
+		\noalign{\smallskip}
+			5	& Simple &	$>$16057(1) &	16758 &	15920 \\
+				& Hard &	$>$217215(16) &	$>$269064(4) & $>$148342(4) \\
+		\noalign{\smallskip}\hline
 	\end{tabular}
-    \caption{Average number of iterations}
-    \label{table:average_iters}
 \end{table}
@@
 \begin{table}
-                \caption{Results on CPU}
+                \caption{Speedup on CPU}
                 \label{table:average_AS_CPU}
@@
                 \begin{tabular}{cccccccccccc}
-                               \hline\noalign{\smallskip}
-                & \multicolumn{5}{c}{ Acceleration of iterations} & & \multicolumn{5}{c}{ Time acceleration}\\
+                               %\hline\noalign{\smallskip}
+																	 & \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
                                \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
-                               $p$ & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
+                                   & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
                                \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
@@
 
-   Let us now evaluate the acceleration achieved using parallel GSA using an adaptive dimensionality reduction scheme based on the number $P$ of cores used.
-Table \ref{table:average_AS_CPU} shows the acceleration by iterations and the time acceleration obtained when solving a series of problems on the CPU, compared to GSA.
+   Let us now evaluate the acceleration achieved using parallel GSA using an adaptive dimensionality reduction scheme based on the number $p$ of cores used.
+Table \ref{table:average_AS_CPU} shows the speedup obtained when solving a series of problems on the CPU, compared to GSA.
 
 \begin{table}
-                \caption{Results on CPU}
+                \caption{Speedup on CPU}
                 \label{table:average_MMAS_CPU}
@@
 
-   Table \ref{table:average_MMAS_CPU} shows the acceleration in iterations and  the acceleration in time of the algorithm that combines multiple sweeps and an adaptive scheme for solving a series of problems on the CPU, compared to the sequential launch of the GSA.
-Two scans and, accordingly, two processes were used; each process used $P$ threads, and calculations were performed on a single cluster node.
+   Table \ref{table:average_MMAS_CPU} shows the speedup of the algorithm that combines multiple evolvents and an adaptive scheme for solving a series of problems on CPU, compared to the sequential launch of the GSA.
+Two evolvents and, accordingly, two processes were used; each process used $p$ threads, and calculations were performed on a single cluster node.
 
@@
                 \begin{tabular}{cccccccccccc}
-                               \hline\noalign{\smallskip}
-                & \multicolumn{5}{c}{ Acceleration of iterations} & & \multicolumn{5}{c}{ Time acceleration}\\
+                               %\hline\noalign{\smallskip}
+																& \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
                                \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
-                               $p$ & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
+                               & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
                                \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
@@
 \begin{table}
-                \caption{Results on GPU}
+                \caption{Speedup on GPU}
                 \label{table:average_MMAS_GPU}
@@
                 \begin{tabular}{cccccccccccc}
-                               \hline\noalign{\smallskip}
-                & \multicolumn{5}{c}{ Acceleration of iterations} & & \multicolumn{5}{c}{ Time acceleration}\\
+                               %\hline\noalign{\smallskip}
+                & \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
                                \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
-                               $p$ & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
+                               & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
                                \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
@@
 
-Tables \ref{table:average_AS_GPU} show the acceleration obtained when solving a series of problems on the GPU using an adaptive scheme compared to a similar launch on the CPU using 4 threads.
-Tables \ref{table:average_MMAS_GPU} show the acceleration of the algorithm combining multiple sweeps and an adaptive circuit when solving a series of problems on the GPU, compared to an adaptive scheme on the CPU using 4 threads.
-Two sweeps and, accordingly, two processes were used; each process used $P$ threads, and calculations were performed on a single cluster node.
+Tables \ref{table:average_AS_GPU} show the speedup obtained when solving a series of problems on the GPU using an adaptive scheme compared to a similar launch on CPU using 4 threads.
+Tables \ref{table:average_MMAS_GPU} show the speedup of the algorithm combining multiple evolvents and an adaptive scheme when solving a series of problems on GPU, compared to an adaptive scheme on CPU using 4 threads.
+Two evolvents and, accordingly, two processes were used; each process used $p$ threads, and calculations were performed on a single cluster node.
 
@@
     \centering
+    \caption{Speedup on GPU}
+    \label{table:GKLS6_average_MMAS_GPU}
     \begin{tabular}{|p{4cm}|c|c|c|}
     \hline
-      & P=2 & P=4 & P=8 \\ \hline
-    Acceleration of iterations&   30,79 & 92,68 & 596,97  \\ \hline
-    Time acceleration         &   1,86  & 1,49  & 2,45    \\ \hline
+      & p=2 & p=4 & p=8 \\ \hline
+    Iterations &   30,79 & 92,68 & 596,97  \\ \hline
+    Time  &   1,86  & 1,49  & 2,45    \\ \hline
 
     \end{tabular}
-    \caption{Acceleration of iterations and time on the GPU}
-    \label{table:GKLS6_average_MMAS_GPU}
 \end{table}

### 2020-04-14 2020/RSD_GPU/RSD2020.tex
Поправил вид последней таблицы. Заменил картинки на pdf.
Вернул закомментированное
one more reference

@@
 Thus, calculating the optimization criterion can be implemented on the GPU, and the role of the optimization algorithm (running on the CPU) consists in the effective selection of points for conducting parallel trials.
-This scheme of working with the accelerator is fully consistent with the work of the parallel global search algorithm developed at the Lobachevsky State University of Nizhni Novgorod and presented in a series of publications \cite{Barkalov2014,Barkalov2020,Barkalov2019,Grishagin2016,Grishagin2015,Grishagin2018,Strongin2000}.
+This scheme of working with the accelerator is fully consistent with the work of the parallel global search algorithm developed at the Lobachevsky State University of Nizhni Novgorod and presented in a series of publications \cite{Barkalov2014,Barkalov2020,Barkalov2019,Grishagin2016,Grishagin2015,Grishagin2018,globalizerSystem}.
 
@@
 \end{displaymath}
-
 For the rationale in using this method of organizing parallel computing see \cite{Strongin2000}. 
@@
     \centering
-		\includegraphics[width=1.0\textwidth]{GPU_Algorithm.eps}
+		\includegraphics[width=0.9\textwidth]{GPU_Algorithm.pdf}
 		\caption{Scheme of information exchanges in the GPU algorithm}\label{fig:GPU_Algorithm}
@@
 We take a small number of nesting levels, 2 to 3, in which the original large dimensionality problem is divided into 2-3 nested lower dimensional subproblems.
-We will use multiple sweeps only at the upper level of nesting, corresponding to the u1 variable block.
+We will use multiple evolvents only at the upper level of nesting, corresponding to the u1 variable block.
 This subproblem will be reduced to a set of one-dimensional problems that will be solved in parallel, each in a separate process.
@@
     \centering
-		\includegraphics[width=1.0\textwidth]{Parallel_Computing.eps}
+		\includegraphics[width=0.7\textwidth]{Parallel_Computing.pdf}
 		\caption{Diagram of parallel computing on a cluster}\label{fig:Parallel_Computing}
@@
                 \center
-                \begin{tabular}{cccccccccccc}
+                \begin{tabular}{lccccccccccc}
                                %\hline\noalign{\smallskip}
@@
 \begin{table}
-    \centering
                 \caption{Speedup on GPU}
-    \label{table:GKLS6_average_MMAS_GPU}
-    \begin{tabular}{|p{4cm}|c|c|c|}
-    \hline
-      & p=2 & p=4 & p=8 \\ \hline
-    Iterations &   30,79 & 92,68 & 596,97  \\ \hline
-    Time  &   1,86  & 1,49  & 2,45    \\ \hline
+                \label{table:average_MMAS_GPU}
+                \center
+                \begin{tabular}{lcc}
+                               %\hline\noalign{\smallskip}
+                &  Iteration speedup &   Time speedup\\
 
+                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
+                               $p=64$  &    11,32	 & 6,83	  \\
+                               $p=256$ &    38,96	 & 31,52  \\
+							   $p=1024$&    128,39	 & 83,46  \\
+                               \noalign{\smallskip}\hline
                 \end{tabular}

### 2020-04-15 2020/RSD_GPU/RSD2020.tex
Delete one table

@@
 %
-\author{Konstantin Barkalov\letter%\orcidID{0000-0001-5273-2471} 
-\and Ilya Lebedev%\orcidID{0000-0002-8736-0652}
+\author{Konstantin Barkalov%\orcidID{0000-0001-5273-2471} 
+\and Ilya Lebedev\letter%\orcidID{0000-0002-8736-0652}
 }
@@
 
-The considered algorithm for solving this problem involves constructing a sequence of points \(x_k\), in which the values of objective function \(z_k = f(x_k)\) are calculated.
+The algorithm for solving this problem (Global Search Algorithm, GSA) involves constructing a sequence of points \(x_k\), in which the values of objective function \(z_k = f(x_k)\) are calculated.
 We will call the process of computing the value of a function at a single point a trial.
 Assume that we have \(p\geqslant 1\) computational elements at our disposal and \(p\) trials which are performed simultaneously (synchronously) within a single iteration of the method.
-Let \(k(n)\) denote the total number of trials performed, after \(n\) parallel iterations.
+Let \(k(n)\) denote the total number of trials performed after \(n\) parallel iterations.
 
-In the first iteration of the method, the trial is carried out at an arbitrary internal point \(x^1\) of the interval \([0,1]\).
+At the first iteration of the method, the trial is carried out at an arbitrary internal point \(x^1\) of the interval \([0,1]\).
 Let \(n>1\) iterations of the method be performed, during which trials were carried out at \(k = k(n)\) points \(x^i, 1\leqslant i\leqslant k\).
@@
 
-Step 2. Assuming \(z_i=f(y(x_i)),1\leqslant i\leqslant k\) calculate the values
+Step 2. Assuming \(z_i=f(y(x_i)),1\leqslant i\leqslant k\), calculate the values
 \begin{displaymath}
 \label{step2}
-\mu=\max_{1\leqslant i\leqslant k}\dfrac{|z_i-z_{i-1}|}{\Delta_i},
+\mu=\max_{1\leqslant i\leqslant k}\dfrac{|z_i-z_{i-1}|}{\Delta_i}, \;
 \begin{matrix}
@@
     \begin{matrix}
-    r\mu,\mu>0 \\
-    1,\mu=0
+    r\mu,\mu>0, \\
+    1,\mu=0,
     \end{matrix} \right.
@@
 \begin{displaymath}
-f_{k}^{*}=\underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}}),   x_{k}^{*}=\arg \quad \underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}}).
+f_{k}^{*}=\underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}}),   x_{k}^{*}=\arg \; \underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}}).
 \end{displaymath}
@@
 However, some information about the proximity of points in multidimensional space is lost, since the point \(x\in [0,1]\) has only left and right neighbors, and the corresponding point \(y(x) \in R^N\) has neighbors in \(2^N\) directions.
-As a result, when using evolvents, the images \(y' , y''\) that are close in \(N\)-dimensional space can correspond to rather distant prototypes \(x' , x''\) on the interval \([0,1]\).
+As a result, when using evolvents, the images \(y' , y''\) that are close in \(N\)-dimensional space can correspond to rather distant preimages \(x' , x''\) on the interval \([0,1]\).
 This property leads to redundant calculations, because several limit points \(x' , x''\) of the test sequence generated by the method on the segment \([0,1]\), can correspond to a single limit point \(y\) in \(N\)-dimensional space.
@@
 instead of using a single Peano curve \(y(x)\) (see \cite{Strongin1992,Strongin2000}).
-For example, each Peano curve \(y^i(x)\) from \(Y_L(x)\) can be obtained as a result of some shift along the main diagonal of the hyperinterval \(D\), the original curve \(y(x)\).
-Another way is to rotate the evolvent around the origin.
-The set of evolvents that have been constructed allows us to obtain for any close images \(y', y''\), close preimages \(x', x''\) for some mapping \(y^i(x)\).
+For example, each Peano curve \(y^i(x)\) from \(Y_L(x)\) can be obtained as a result of some shifting \(y(x)\) along the main diagonal of the hyperinterval \(D\).
+Another way is to rotate the evolvent \(y(x)\) around the origin.
+The set of evolvents that have been constructed allows us to obtain for any close images \(y', y''\) close preimages \(x', x''\) for some mapping \(y^i(x)\).
 
@@
 
-In relation to global optimization methods, an operation that can be effectively implemented on the GPU is the parallel calculation of many values of the objective function at once.
-Naturally, this requires implementing a procedure for calculating the value of a function on the GPU.
-Data transfers from the CPU to the GPU will be minimal: one need only transfer the coordinates of the trial points to the GPU, and get back the function values at these points.
-Functions that determine the processing of trial results in accordance with the algorithm, and require work with large amounts of accumulated search information, can be effectively implemented on the CPU.
+In relation to global optimization methods, an operation that can be efficiently implemented on GPU is the parallel calculation of many values of the objective function at once.
+Naturally, this requires implementing a procedure for calculating the value of a function on GPU.
+Data transfers from CPU to GPU will be minimal: one need only transfer the coordinates of the trial points to the GPU, and get back the function values at these points.
+Functions that determine the processing of trial results in accordance with the algorithm, and require work with large amounts of accumulated search information, can be effectively implemented on CPU.
 
-   The general scheme for organizing calculations using the GPU is shown in \ref{fig:GPU_Algorithm}.
-In accordance with this scheme, steps 1--4 of the parallel global search algorithm are performed on the CPU.
-The coordinates of the $p$ trial points calculated in step 4 of the algorithm are accumulated in the intermediate buffer and then transmitted to the GPU.
-On the GPU, the function values are calculated at these points, after which the trial results (again through the intermediate buffer) are transferred to the CPU.
+   The general scheme for organizing calculations using GPU is shown in Fig.
+\ref{fig:GPU_Algorithm}.
+In accordance with this scheme, steps 1--4 of the parallel global search algorithm are performed on CPU.
+The coordinates of the $p$ trial points calculated in step 4 of the algorithm are accumulated in the intermediate buffer and then transmitted to GPU.
+On GPU the function values are calculated at these points, after which the trial results (again through the intermediate buffer) are transferred to CPU.
 
@@
  differs from all the previous subproblems; it calculates the values of the objective function, since ${{\varphi }_{M}}({{u}_{1}},...,{{u}_{M}})=\varphi ({{y}_{1}},...,{{y}_{N}})$.
-This subproblem transfers between the CPU and the GPU.
+This subproblem transfers between CPU and GPU.
 
@@
 
-\begin{table}
-                \caption{Speedup on CPU}
-                \label{table:average_AS_CPU}
-                \center
-                \begin{tabular}{lccccccccccc}
-                               %\hline\noalign{\smallskip}
-																	 & \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
-                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
-                                   & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
-                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
-                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
-                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
-                               $p=2$&   3,49 & 2,35 &   & 0,86 & 2,03 &   & 1,71 & 1,17 &   & 0,95 & 0,4 \\
-                               $p=4$&   8,22 & 7,65 &   & 6,74 & 8,56 &   & 2,01 & 1,9  &   & 1,85 & 1,53 \\
-                               \noalign{\smallskip}\hline
-                \end{tabular}
-\end{table}
+%\begin{table}
+                %\caption{Speedup on CPU}
+                %\label{table:average_AS_CPU}
+                %\center
+                %\begin{tabular}{lccccccccccc}
+                               %%\hline\noalign{\smallskip}
+																	 %& \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
+                               %\noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
+                                   %& \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
+                               %\noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
+                               %& Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
+                               %\noalign{\smallskip} \hline \noalign{\smallskip}                               
+                               %$p=2$&   3,49 & 2,35 &   & 0,86 & 2,03 &   & 1,71 & 1,17 &   & 0,95 & 0,4 \\
+                               %$p=4$&   8,22 & 7,65 &   & 6,74 & 8,56 &   & 2,01 & 1,9  &   & 1,85 & 1,53 \\
+                               %\noalign{\smallskip}\hline
+                %\end{tabular}
+%\end{table}
 
 Let us now evaluate the acceleration achieved using parallel GSA using an adaptive dimensionality reduction scheme based on the number $p$ of cores used.
-Table \ref{table:average_AS_CPU} shows the speedup obtained when solving a series of problems on the CPU, compared to GSA.
+Table \ref{table:average_MMAS_CPU} shows the speedup of the algorithm that combines multiple evolvents and an adaptive scheme for solving a series of problems on CPU, compared to the sequential launch of the GSA.
+Two evolvents and, accordingly, two processes were used; each process used $p$ threads, and calculations were performed on a single cluster node.
+% Table \ref{table:average_AS_CPU} shows the speedup obtained when solving a series of problems on the CPU, compared to GSA.
 
@@
 \end{table}
-
-   Table \ref{table:average_MMAS_CPU} shows the speedup of the algorithm that combines multiple evolvents and an adaptive scheme for solving a series of problems on CPU, compared to the sequential launch of the GSA.
-Two evolvents and, accordingly, two processes were used; each process used $p$ threads, and calculations were performed on a single cluster node.
 
@@
 
-Tables \ref{table:average_AS_GPU} show the speedup obtained when solving a series of problems on the GPU using an adaptive scheme compared to a similar launch on CPU using 4 threads.
-Tables \ref{table:average_MMAS_GPU} show the speedup of the algorithm combining multiple evolvents and an adaptive scheme when solving a series of problems on GPU, compared to an adaptive scheme on CPU using 4 threads.
+Table \ref{table:average_AS_GPU} shows the speedup obtained when solving a series of problems on the GPU using an adaptive scheme compared to a similar launch on CPU using 4 threads.
+Table \ref{table:average_MMAS_GPU} shows the speedup of the algorithm combining multiple evolvents and an adaptive scheme when solving a series of problems on GPU, compared to an adaptive scheme on CPU using 4 threads.
 Two evolvents and, accordingly, two processes were used; each process used $p$ threads, and calculations were performed on a single cluster node.
@@
                 \caption{Speedup on GPU}
-                \label{table:average_MMAS_GPU}
+                \label{table:GKLS6_average_MMAS_GPU}
                 \center

### 2020-04-15 2020/RSD_GPU/RSD2020.tex
Add comment for the last table

@@
 The general assumption is made about the function being optimized: it satisfies the Lipschitz condition with an a priori unknown constant.
-Many approaches to solving problems of this class are based on reducing the dimension of the problem; i.e., addressing a multidimensional problem by solving a family of problems with lower dimension.
+Many approaches to solving problems of this class are based on reducing the dimension of the problem; i.e.
+addressing a multidimensional problem by solving a family of problems with lower dimension.
 In this work, an adaptive dimensionality reduction scheme is investigated, and its implementation using graphic accelerators is proposed.
@@
 However, the potential for graphics accelerators to solve global optimization problems has not yet been fully realized.
-Using the GPUs, they basically parallelize nature-inspired optimization algorithms, which are somehow based on the idea of random search (see, for example, \cite{Ferreiro2013,Garcia2014,Langdon2011}).
+Using  GPUs, they basically parallelize nature-inspired optimization algorithms, which are somehow based on the idea of random search (see, for example, \cite{Ferreiro2013,Garcia2014,Langdon2011}).
 By virtue of their stochastic nature, algorithms of this type guarantee convergence to the global minimum only in the sense of probability, which differentiates them unfavorably from deterministic methods.
@@
 With regard to many deterministic algorithms of Lipschitzian global optimization with guaranteed convergence, parallel variants have been proposed \cite{Evtushenko2009,He2008,Paulavicius2011}.
-However, these versions of algorithms are parallelized on the CPU using shared and/or distributed memory; presently, no GPU implementations have been made.
+However, these versions of algorithms are parallelized on CPU using shared and/or distributed memory; presently, no GPU implementations have been made.
 For example, \cite{Paulavicius2011} describes parallelization of an algorithm based on the ideas of the branch and boundary method using MPI and OpenMP.
@@
 Currently, graphics accelerators can be used to solve problems of this type.
-Moreover, an accelerator can solve several such problems at once \cite{Kindratenko2014}; i.e., using the GPU, one can calculate multiple function values simultaneously.
+Moreover, an accelerator can solve several such problems at once \cite{Kindratenko2014}; i.e., using GPU, one can calculate multiple function values simultaneously.
 
-Thus, calculating the optimization criterion can be implemented on the GPU, and the role of the optimization algorithm (running on the CPU) consists in the effective selection of points for conducting parallel trials.
+Thus, calculating the optimization criterion can be implemented on GPU, and the role of the optimization algorithm (running on CPU) consists in the effective selection of points for conducting parallel trials.
 This scheme of working with the accelerator is fully consistent with the work of the parallel global search algorithm developed at the Lobachevsky State University of Nizhni Novgorod and presented in a series of publications \cite{Barkalov2014,Barkalov2020,Barkalov2019,Grishagin2016,Grishagin2015,Grishagin2018,globalizerSystem}.
@@
 Naturally, this requires implementing a procedure for calculating the value of a function on GPU.
-Data transfers from CPU to GPU will be minimal: one need only transfer the coordinates of the trial points to the GPU, and get back the function values at these points.
+Data transfers from CPU to GPU will be minimal: one need only transfer the coordinates of the trial points to GPU, and get back the function values at these points.
 Functions that determine the processing of trial results in accordance with the algorithm, and require work with large amounts of accumulated search information, can be effectively implemented on CPU.
@@
 \ref{fig:Parallel_Computing} shows the general scheme of organizing calculations using several cluster nodes and several GPUs.
-In accordance with this scheme, nested subproblems  ${{\varphi }_{i}}({{u}_{1}},...,{{u}_{i}})=\underset{{{u}_{i+1}}\in {{D}_{i+1}}}{\mathop{\min }}\,{{\varphi }_{i+1}}({{u}_{1}},...,{{u}_{i}},{{u}_{i+1}})$  with \(i=1,\ldots ,M-2\) are solved using only the CPU.
+In accordance with this scheme, nested subproblems  ${{\varphi }_{i}}({{u}_{1}},...,{{u}_{i}})=\underset{{{u}_{i+1}}\in {{D}_{i+1}}}{\mathop{\min }}\,{{\varphi }_{i+1}}({{u}_{1}},...,{{u}_{i}},{{u}_{i+1}})$  with \(i=1,\ldots ,M-2\) are solved using CPU only.
 The values of the optimized function are not calculated directly in these subproblems: the calculation the function value  ${{\varphi }_{i}}({{u}_{1}},...,{{u}_{i}})$ is a solution to the minimization problem at the next level.
@@
 The numerical experiments were carried out using the Lomonosov supercomputer (Lomonosov Moscow State University).
-Each supercomputer node included two quad-core processors, Intel Xeon X5570 2.93GHz and 12 Gb RAM.
+Each supercomputer node included two quad-core processors Intel Xeon X5570, two NVIDIA Tesla X2070 and 12 Gb RAM.
 When building the Globalizer system for running on the Lomonosov supercomputer, the GCC 4.3.0 compiler, CUDA 6.5 and Intel MPI 2017 were used.
@@
 Note that well-known test problems from the field of multidimensional global optimization are characterized by a short time span for calculating the values of the objective function.
-Therefore, in order to simulate the computational complexity inherent in applied optimization problems, the calculation of the objective function in all experiments was complicated by additional calculations that do not change the form of the function and the location of its minima (summing the segment of the Taylor series).
+Therefore, in order to simulate the computational complexity inherent in applied optimization problems \cite{Modorskii2016}, the calculation of the objective function in all experiments was complicated by additional calculations that do not change the form of the function and the location of its minima (summing the segment of the Taylor series).
 
@@
 Let us now evaluate the acceleration achieved using parallel GSA using an adaptive dimensionality reduction scheme based on the number $p$ of cores used.
-Table \ref{table:average_MMAS_CPU} shows the speedup of the algorithm that combines multiple evolvents and an adaptive scheme for solving a series of problems on CPU, compared to the sequential launch of the GSA.
+Table \ref{table:average_MMAS_CPU} shows the speedup of the algorithm that combines multiple evolvents and an adaptive scheme for solving a series of problems on CPU, compared to the sequential launch of the GSA method.
 Two evolvents and, accordingly, two processes were used; each process used $p$ threads, and calculations were performed on a single cluster node.
@@
 \begin{table}
-                \caption{Results on GPU}
+                \caption{Speedup on one GPU}
                 \label{table:average_AS_GPU}
@@
 \begin{table}
-                \caption{Speedup on GPU}
+                \caption{Speedup on two GPUs}
                 \label{table:average_MMAS_GPU}
@@
 
-Table \ref{table:average_AS_GPU} shows the speedup obtained when solving a series of problems on the GPU using an adaptive scheme compared to a similar launch on CPU using 4 threads.
-Table \ref{table:average_MMAS_GPU} shows the speedup of the algorithm combining multiple evolvents and an adaptive scheme when solving a series of problems on GPU, compared to an adaptive scheme on CPU using 4 threads.
-Two evolvents and, accordingly, two processes were used; each process used $p$ threads, and calculations were performed on a single cluster node.
+Table \ref{table:average_AS_GPU} shows the speedup obtained when solving a series of problems on one GPU using an adaptive scheme compared to a similar launch on CPU using 4 threads.
+Table \ref{table:average_MMAS_GPU} shows the speedup of the algorithm combining multiple evolvents and an adaptive scheme when solving a series of problems on two GPUs compared to an adaptive scheme on CPU using 4 threads.
+Two evolvents and, accordingly, two processes were used; each process used $p$ threads on each GPU; all computations were performed on a single cluster node.
 
 \begin{table}
-                \caption{Speedup on GPU}
+                \caption{Speedup on six GPUs}
                 \label{table:GKLS6_average_MMAS_GPU}
                 \center
-                \begin{tabular}{lcc}
+                \begin{tabular}{lccc}
                                %\hline\noalign{\smallskip}
-                &  Iteration speedup &   Time speedup\\
-
+                &  Iteration speedup & &   Time speedup \\
                                \noalign{\smallskip} \hline \noalign{\smallskip}                               
-                               $p=64$  &    11,32	 & 6,83	  \\
-                               $p=256$ &    38,96	 & 31,52  \\
-							   $p=1024$&    128,39	 & 83,46  \\
+                               $p=64$  &    11,32	 & & 6,83	  \\
+                               $p=256$ &    38,96	 & & 31,52  \\
+															 $p=1024$&    128,39	 & & 83,46  \\
                                \noalign{\smallskip}\hline
@@
 
-%Следующий эксперимент будет проведен на 6-ти мерных задачах GKLS.
+%Table \ref{table:average_AS_GPU} shows the speedup obtained when solving a series of problems on one GPU using an adaptive scheme compared to a similar launch on CPU using 4 threads.
+Table \ref{table:average_MMAS_GPU} shows the speedup of the algorithm combining multiple evolvents and an adaptive scheme when solving a series of problems on two GPUs compared to an adaptive scheme on CPU using 4 threads.
+Two evolvents and, accordingly, two processes were used; each process used $p$ threads on each GPU; all computations were performed on a single cluster node.
+
+The last series of experiments has been carried out on 100 six-dimensional problems from the GKLS Simple class. 
+Table \ref{table:GKLS6_average_MMAS_GPU} shows the speedup of the algorithm combining multiple evolvents and the adaptive scheme when solving the problems on 3 cluster nodes (using 2 GPUs per node, 6144 GPU threads in all) compared to the adaptive scheme on CPU using 4 threads.
+
+%Заключительный эксперимент был проведен на 6-ти мерных задачах класса GKLS Simple.
 В таблице \ref{table:GKLS6_average_MMAS_GPU} приведено ускорение по времени и по итерациям полученное при решении серии задач на GPU с использованием алгоритма объединяющего множественные развертки и адаптивную схему по сравнению с адаптивной схемой на CPU использующей 4 потока.
-Эксперимент проводился на трех узлах кластера и спользовалось до 6144 нитей.
+Эксперимент проводился на трех узлах кластера, шесть разверток = шесть проецессов, каждый процесс - p нитей.
+Максимальное число нитей 6144.
 
@@
 
-In this paper, we consider a parallel algorithm for solving multidimensional multiextremal optimization problems and its implementation on the GPU.
+In this paper, we consider a parallel algorithm for solving multidimensional multiextremal optimization problems and its implementation on GPU.
 In order to experimentally confirm the theoretical properties of the parallel algorithm under consideration, computational experiments were carried out on a series of several hundred test problems of different dimensions.
-The parallel algorithm demonstrates good acceleration, both in the GPU and CPU versions.
+The parallel algorithm demonstrates good speedup, both in the GPU and CPU versions.
 

### 2020-04-15 2020/RSD_GPU/RSD2020.tex
Поправлены таблицы

@@
 
-\section{Multidimensional parallel global search algorithm}
+\section{Multidimensional parallel global search algorithm}\label{Sec_GSA}
 Consider the problem of finding the global minimum of the \(N\)-dimensional function \(\varphi(y)\) in the hyperinterval  \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
@@
 		\item to calculate the value of the function of the \(i\)-th level from (\ref{subTasks}), a new \(i+1\)-th level problem is generated in which trials are carried out, after which a new generated problem is included in the set of existing problems to be solved;
-		\item the iteration of the global search consists in choosing \(p\) (the most promising) problem from the set of existing problems in which trials are carried out; points for new trials are determined in accordance with the algorithm of subsection 1;
+		\item the iteration of the global search consists in choosing \(p\) (the most promising) problem from the set of existing problems in which trials are carried out; points for new trials are determined in accordance with the parallel global search algorithm from Section \ref{Sec_GSA};
 		\item the minimum values of functions from (\ref{subTasks}) are their current estimates based on the accumulated search information.
@@
 A brief description of the main steps of a block adaptive dimensionality reduction scheme is as follows.
-Let the nested subproblems in the form (\ref{subTasks}) be solved using the global search algorithm described in Section 2.
+Let the nested subproblems in the form (\ref{subTasks}) be solved using the global search algorithm described in Section \ref{Sec_GSA}.
 Then each subproblem (\ref{subTasks}) can be assigned a numerical value called the characteristic of this problem.
@@
 
-\ref{fig:Parallel_Computing} shows the general scheme of organizing calculations using several cluster nodes and several GPUs.
+Fig.
+\ref{fig:Parallel_Computing} shows the general scheme of organizing computations using several cluster nodes and several GPUs.
 In accordance with this scheme, nested subproblems  ${{\varphi }_{i}}({{u}_{1}},...,{{u}_{i}})=\underset{{{u}_{i+1}}\in {{D}_{i+1}}}{\mathop{\min }}\,{{\varphi }_{i+1}}({{u}_{1}},...,{{u}_{i}},{{u}_{i+1}})$  with \(i=1,\ldots ,M-2\) are solved using CPU only.
@@
 
-Note that well-known test problems from the field of multidimensional global optimization are characterized by a short time span for calculating the values of the objective function.
+Note that well-known test problems from the field of multidimensional global optimization are characterized by a short time of calculating the values of the objective function.
 Therefore, in order to simulate the computational complexity inherent in applied optimization problems \cite{Modorskii2016}, the calculation of the objective function in all experiments was complicated by additional calculations that do not change the form of the function and the location of its minima (summing the segment of the Taylor series).
@@
 A numerical comparison was carried out on the Simple and Hard function classes of dimension 4 and 5 from \cite{Gaviano2003}.
-The global minimum $y*$ was considered found if the algorithm generated a trial point $y^k$ in the $\delta$-neighborhood of the global minimum, i.e.
+The global minimum $y^*$ was considered found if the algorithm generated a trial point $y^k$ in the $\delta$-neighborhood of the global minimum, i.e.
 $\left\| {{y}^{k}}-{{y}^{*}} \right\| \leqslant \delta $.
-Also, the size of the neighborhood was chosen (in accordance with \cite{Sergeyev2006}) as $\delta =\left\| b-a \right\|\sqrt[N]{\Delta }$, here $N$ is the dimension of the problem to be solved , a and b are the boundaries  of the search domain $D$, the parameter  $\Delta ={{10}^{-6}}$ for N=4 and $\Delta ={{10}^{-7}}$ for $N=5$.
-When using the GSA method for the Simple class, the parameter $r=4.5$ was selected, for the Hard class $r=5.6$; the parameter for constructing the Peano curve was fixed, $m=10$.
+The size of the neighborhood was chosen (in accordance with \cite{Sergeyev2006}) as $\delta =\left\| b-a \right\|\sqrt[N]{\Delta }$, here $N$ is the dimension of the problem to be solved , $a$ and $b$ are the boundaries  of the search domain $D$, the parameter  $\Delta ={{10}^{-6}}$ for $N=4$ and $\Delta ={{10}^{-7}}$ for $N=5$.
+When using the GSA method for the Simple class, the parameter $r=4.5$ was selected, for the Hard class $r=5.6$; the parameter for constructing the Peano curve was $m=10$.
 The maximum number of iterations allowed was $K_{max} = 10^6$.
@@
 
-As can be seen from Table \ref{table:average_iters}, the sequential GSA surpasses the DIRECT and DIRECT\textit{l} methods in all classes of problems in terms of the average number of iterations.
+As can be seen from Table \ref{table:average_iters}, the sequential GSA surpasses DIRECT and DIRECT\textit{l} methods in all classes of problems in terms of the average number of iterations.
 At the same time, in the 5-Hard class, none of the methods solved all the problems: DIRECT failed to solve 16 problems, DIRECTl and GSA -- 4 problems each.
@@
                                %\noalign{\smallskip} \hline \noalign{\smallskip}                               
-                               %$p=2$&   3,49 & 2,35 &   & 0,86 & 2,03 &   & 1,71 & 1,17 &   & 0,95 & 0,4 \\
-                               %$p=4$&   8,22 & 7,65 &   & 6,74 & 8,56 &   & 2,01 & 1,9  &   & 1,85 & 1,53 \\
+                               %$p=2$&   3,5 & 2,4 &   & 0,9 & 2,0 &   & 1,7 & 1,2 &   & 1,0 & 0,4 \\
+                               %$p=4$&   8,2 & 7,7 &   & 6,7 & 8,6 &   & 2,0 & 1,9  &   & 1,9 & 1,5 \\
                                %\noalign{\smallskip}\hline
@@
                                \noalign{\smallskip} \hline \noalign{\smallskip}                               
-                               $p=2$&    6,62  & 3,23  &   & 2,13 & 6,07  &   & 3,22 & 1,48 &   & 0,66 & 1,93 \\
-                               $p=4$&    21,46 &  9,97 &   & 6,9  & 16,27 &   & 5,23 & 2,42 &   & 0,85 & 1,9 \\
+                               $p=2$&    6,6  & 3,2   &   & 2,1  & 6,6  &   & 3,2 & 1,5 &   & 0,7 & 2,1 \\
+                               $p=4$&    21,5 &  10,0 &   & 6,9  & 19,2 &   & 5,2 & 2,4 &   & 0,9 & 2,3 \\
                                \noalign{\smallskip}\hline
@@
                                \noalign{\smallskip} \hline \noalign{\smallskip}                               
-                               $p=64$  &    5,06  & 3,93  &   & 1,19 	 & 5,41  &   & 2,2	    & 1,92	 &   & 0,47	& 2,21 \\
-                               $p=256$ &    19,91 & 14,95 &   & 11,9 	 & 23,36 &   & 8,28	& 6,91	     &   & 3,4	    & 6,1 \\
-							   $p=1024$&    15,62 & 52,91 &   & 22,12	 & 55,22 &   & 5,21	& 19,95      &   & 2,18	& 4,82 \\
+                               $p=64$  &    5,1  & 3,9  &   & 1,2 	 & 9,3   &   & 2,2	& 1,9	     &   & 0,5	& 4,0  \\
+                               $p=256$ &    19,9 & 15,0 &   & 11,9 	 & 39,6  &   & 8,3	& 6,9	     &   & 3,4	& 11,0 \\
+							   $p=1024$&    15,6 & 52,9 &   & 22,1	 & 105,7 &   & 5,2	& 20,0      &   & 2,2	& 10,2 \\
                                \noalign{\smallskip}\hline
@@
                                \noalign{\smallskip} \hline \noalign{\smallskip}                               
-                               $p=64$  &    11,32	 & 6,83	 &   & 8,19	     & 10,66  &   & 2,44	 & 1,63	 &   & 1,33	 & 1,58 \\
-                               $p=256$ &    38,96	 & 31,52 &   & 25,43	 & 48,23  &   & 8,01	 & 7,04	 &   & 2,31	 & 3,42 \\
-							   $p=1024$&    128,39	 & 83,46 &   & 98,43	 & 146,23 &   & 20,45	 & 11,06 &   & 2,75	 & 2,48 \\
+                               $p=64$  &    11,3	 & 6,8	 &   & 8,2	 & 19,8  &   & 2,4	 & 1,6	 &   & 1,3	 & 3,2 \\
+                               $p=256$ &    39,0	 & 31,5 &   & 25,4	 & 57,1  &   & 8,0	 & 7,0	 &   & 2,3	 & 2,4 \\
+							   $p=1024$&    128,4	 & 83,5 &   & 98,4	 & 267,9 &   & 20,4  & 11,0  &   & 2,8	 & 5,2 \\
                                \noalign{\smallskip}\hline
@@
                                \noalign{\smallskip} \hline \noalign{\smallskip}                               
-                               $p=64$  &    11,32	 & & 6,83	  \\
-                               $p=256$ &    38,96	 & & 31,52  \\
-															 $p=1024$&    128,39	 & & 83,46  \\
+                               $p=64$  &    11,3	 & & 6,8	  \\
+                               $p=256$ &    39,0	 & & 31,5  \\
+															 $p=1024$&    128,4	 & & 83,5  \\
                                \noalign{\smallskip}\hline

### 2020-05-01 2019/NUMTA_Barkalov/SoftComputing/article.tex
remove coloring

@@
 \usepackage[misc]{ifsym}
-\usepackage{xcolor}
+%\usepackage{xcolor}
 % etc.
@@
 
-{ \color{red}
 	2. If $\nu(x_{i-1}) = \nu(x_i) = M$, then, taking into account (\ref{current_lower_bounds}), the following inequality will be valid
@@
 
-}	
 %\begin{enumerate} 
@@
 
-{ \color{blue}
-
 	This statement can be interpreted as the incomparability of the local and global characteristics: the global characteristic of the best interval is larger if $\nu(x_{i-1})=\nu(x_i)=M$.
@@
 Based on the relations (\ref{R_glob_equal_R_loc}) and (\ref{new_rho}), one can formulate an \textit{index algorithm with a dual estimate of the Lipschitz constants} (hereafter referred to as IA-DL).
-
-}
 
@@
 	Let us formulate and prove the convergence conditions for the new algorithm IA-DL.\\
-	{ \color{blue}
 
@@
 The lemma is proved.
-}
 
@@
 	At every iteration of the IA-DL method there will exist such interval $[x_{j-1}, x_j]$, $j=j(k)$, that $\bar x \in [x_{j-1}, x_j]$.
-	{ \color{blue}
 According to our assumption that $\bar y = y(\bar x)$ is not the limit point, there will exist such number $\bar k$, starting from which the trial points will not fall into the interval $[x_{j-1}, x_j]$.
 That is, to conduct a new trial at the point $x^{k+1} \in [x_{t-1}, x_t]$ for each $k\geq \bar k$ such interval $[x_{t-1}, x_t], t=t(k),$ will be selected that $t(k) \neq j(k)$.
-	}
 
@@
 	& + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} = \\
-	& = { \color{red} \Delta_i\left(1-\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} } > 0.
+	& = \Delta_i\left(1-\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
 \end{align*}
@@
 
-{ \color{red}	
 	At the beginning of the proof of the theorem we have assumed that $x'$ is a limit point for the IA-DL and  $x' \neq \bar x$.
@@
 Characteristics $R(t)$ of the intervals of the specified sequence will be positive and $R(t) \rightarrow 0$ at $k \rightarrow \infty$.
-}
 
@@
 \end{equation}
-{\color{blue}
 Indeed, both $j$ and $t$ depend on $k$, as the numbering of the trial points (by the lower index) changes at every iteration and the numbering of the intervals changes respectively.
@@
 %new text end
-}
 

### 2020-05-02 2019/NUMTA_Barkalov/SoftComputing/article.tex
fix some comments of the Reviewer 1

@@
 In particular, this method cannot be applied to the problems with the partially defined functions (i.e.
-when any constraint is violated, the values of all the other functions of the problem remain undefined).
+some of the functions of the problem are defined only at those points which satisfy a selection of constraints).
 
@@
 An underestimate of the true value of the constant may lead to the loss of the algorithm convergence to the global solution.
-At the same time, too high a value of the constant estimate not corresponding to real behavior of the function results in slow convergence of the algorithm to the global minimum point.
+At the same time, a too high value of the constant estimate not corresponding to real behavior of the function results in slow convergence of the algorithm to the global minimum point.
 
@@
 \item global estimate of the constant $L$ in the whole search domain $D$ (\cite{Strongin2000,Pinter1996,Horst1996}).
-\item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ (\cite{Kvasov2003,Sergeyev2010,Sergeyev2016}).
+\item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ (\cite{Sergeyev1998,Kvasov2003,Sergeyev2010,Sergeyev2016}).
 \item choice of the estimates of the constant $L$ from some set of possible values (\cite{Jones2009,Jones1993,Gablonsky2001,Sergeyev2006}).
@@
 	Each of the approaches listed above has its own advantages and drawbacks.
-For example, the use of the global estimate only in the entire search domain may slow down the convergence of the algorithm to the global minimum point.
+For example, the only use of the global estimate in the entire search domain may slow down the convergence of the algorithm to the global minimum point.
 On the other hand, the use of the local estimates accelerating the convergence requires an adequate adjustment of the algorithm parameters in order to ensure the global convergence.
@@
 \end{align}
-with a priori unknown constants $L_j, 1\leq j \leq m+1$, and may be multiextremal.
+with a priori unknown constants $L_j, \; 0<L_j<\infty, \; 1\leq j \leq m+1$, and may be multiextremal.
 
@@
 \normalsize
-and the set of solutions $Y_{\epsilon}$, which are not worse than the $\varepsilon$-reserved solution in the objective function value, i.e.
+where the algorithm parameters $\epsilon_j, \; 1\leq j \leq m,$ are non negative numbers, and the set of solutions $Y_{\epsilon}$, which are not worse than the $\varepsilon$-reserved solution in the objective function value, i.e.
 \small
@@
 
-	The existence of the $\varepsilon$-reserved solutions of problem (\ref{problem}) can be interpreted as some analog of the regularity conditions in classical nonlinear programming problems.
+	The existence of the $\varepsilon$-reserved solutions of problem (\ref{problem}) can be interpreted as some analogous of the regularity conditions in classical nonlinear programming problems.
 The applied role of this condition should be also noted.
@@
 $$
-where the last inequality is insufficient if $\nu = m+1$.
+where the last inequality is not necessary if $\nu = m+1$.
 
@@
 	An approach, in which two estimates of the Lipschitz constant for each function are used in the rules of the algorithm, is promising.
-One of these estimates is significantly greater than the other, which corresponds to the use of two sets of parameters $r_{\nu}^{glob}$ and $r_{\nu}^{loc}$ in the algorithm, where $r_{\nu}^{glob}>r_{\nu}^{loc}>1, 1 \leq \nu \leq m+1$.
+One of these estimates is significantly greater than the other, which corresponds to the use of two sets of parameters $r_{\nu}^{glob}$ and $r_{\nu}^{loc}$ in the algorithm, where $r_{\nu}^{glob}>r_{\nu}^{loc}>1$ and $1 \leq \nu \leq m+1$.
 When using the parameter $r_{\nu}^{loc}$ we shall deal with the smaller estimate of the Lipschitz constant, and using the parameter $r_{\nu}^{glob}$ will correspond to the larger one.
@@
 The best point is marked by red.
-For both methods, the evolvent with the parameter $m=10$ was used, the accuracy of the search for the solution $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ from (\ref{epsilon_nu}).
+For both methods, the evolvent with the parameter $m=10$ was used, the accuracy of the search for the solution was $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ is from (\ref{epsilon_nu}).
 For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$, were used.

### 2020-05-04 2019/NUMTA_Barkalov/SoftComputing/article.tex
New Fig1 (without decsription)
all comment fixed, but not checked

@@
 \usepackage[misc]{ifsym}
-%\usepackage{xcolor}
+\usepackage{xcolor}
 % etc.
@@
 Fig. \ref{fig:eps_reserved_solution} illustrates the concepts introduced.
+{\color{red} The figure shows global minimizer $y^*$, $\epsilon$-reserved solution $y_\epsilon$ and the set $Y_\epsilon$ of all admissible points which are not worse than the $\epsilon$-reserved solution.
+}
 
 \begin{figure}
-	\center{\includegraphics[width=0.6\linewidth]{Fig1}}
-	\caption{$\varepsilon$-reserved solution.}
+	\center{\includegraphics[width=0.6\linewidth]{Fig1.pdf}}
+	\caption{ {\color{red} An $\varepsilon$-reserved solution to a two-dimensional problem. } }
 	\label{fig:eps_reserved_solution}
@@
 This situation is typical for the optimal design problems when some conditions for the functioning of the modeled object are not met and its other characteristics appear to be undefined. 
-In this approach, the first violation of any constraint indicating that the point $y$ being considered is no longer feasible terminates further computations at the point $y$.
+{\color{red} In this approach, the violation of any constraint at the point $y$ terminates the computation of the values of other functions at this point. }
 
@@
 The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000,Sergeyev2013}.
-Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$, where the number $m>1$ is a parameter (the \textit{density} of the evolvent).
+Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-l}$, where the integer number $l>1$ is a parameter (the \textit{density} of the evolvent).
 The evolvent density is determined by the accuracy of the possible practical implementation of the problem solution found.
 For  illustration, Fig.
-\ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
+\ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $l=4,5$, correspondingly.
 \begin{figure}
@@
 	\end{minipage}
-	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
+	\caption{Evolvent for $N=2$ and $l=4$ (a), $l=5$ (b)}
 	\label{fig:evolvent}
@@
 The best point is marked by red.
-For both methods, the evolvent with the parameter $m=10$ was used, the accuracy of the search for the solution was $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ is from (\ref{epsilon_nu}).
+For both methods, the evolvent with the parameter $l=10$ was used, the accuracy of the search for the solution was $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ is from (\ref{epsilon_nu}).
 For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$, were used.
@@
 	\item the global minimizer is known in advance and can be located at the boundary of the feasible domain;
-	\item the number of constraints active in the global minimizer can be predefined.
+	\item {\color{red} the number of constraints active at the global minimizer can be predefined (an inequality constraint is called active at some point if at this point the inequality turns into equality)}.
 \end{itemize}
@@
 $\| y^{\ast} - y^k \| < \varepsilon \|b - a\|$, where $a$ and $b$ are the boundaries of the search domain $D$.
-To reduce the dimensionality, the evolvent constructed with the parameter $m=10$ was used in both algorithms.
+To reduce the dimensionality, the evolvent constructed with the parameter $l=10$ was used in both algorithms.
 The relative accuracy of the search for the solution at $N = 2, 3$ was $\epsilon=0.01$ and for the solution at $N=4,5$  it was $\epsilon=0.03$.

### 2020-05-26 2020/RSD_GPU/RSD2020.tex
Add Toropov as co-author

@@
 %
-\author{Konstantin Barkalov%\orcidID{0000-0001-5273-2471} 
-\and Ilya Lebedev\letter%\orcidID{0000-0002-8736-0652}
+\author{
+Konstantin A. Barkalov%(ORCID 0000-0001-5273-2471)
+$^1$ \and
+Ilya G. Lebedev \letter%(ORCID 0000-0002-8736-0652)
+$^1$ \and 
+Vassili V. Toropov%(ORCID 0000-0002-7017-5055)
+$^{1,2}$
 }
@@
 %
-\authorrunning{Konstantin Barkalov \and Ilya Lebedev} % abbreviated author list (for running head)
+\authorrunning{K.A. Barkalov \and I.G. Lebedev \and V.V. Toropov} % abbreviated author list (for running head)
 %
 %%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{Konstantin Barkalov and Ilya Lebedev}
+%\tocauthor{Konstantin Barkalov and Ilya Lebedev }
 %
-\institute{Lobachevsky State University of Nizhni Novgorod, Russia  \\
+
+\institute{$^1$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
 \email{konstantin.barkalov@itmm.unn.ru},
-	\email{ilya.lebedev@itmm.unn.ru}
+\email{ilya.lebedev@itmm.unn.ru}\\
+$^2$Queen Mary University of London, London, UK\\
+\email{v.v.toropov@qmul.ac.uk}\\
 }

### 2020-05-27 2019/NUMTA_Barkalov/SoftComputing/article.tex

@@
 \usepackage{hyperref,amsmath}
-\usepackage[pagewise]{lineno}
+%\usepackage[pagewise]{lineno}
 \usepackage{multirow}
@@
 \item global estimate of the constant $L$ in the whole search domain $D$ (\cite{Strongin2000,Pinter1996,Horst1996}).
-\item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ (\cite{Sergeyev1998,Kvasov2003,Sergeyev2010,Sergeyev2016}).
+\item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ (\cite{Sergeyev1998,Kvasov2003,Sergeyev2010}, \cite{Sergeyev2016}).
 \item choice of the estimates of the constant $L$ from some set of possible values (\cite{Jones2009,Jones1993,Gablonsky2001,Sergeyev2006}).
@@
 \normalsize
-Fig. \ref{fig:eps_reserved_solution} illustrates the concepts introduced.
-{\color{red} The figure shows global minimizer $y^*$, $\epsilon$-reserved solution $y_\epsilon$ and the set $Y_\epsilon$ of all admissible points which are not worse than the $\epsilon$-reserved solution.
-}
+Fig. \ref{fig:eps_reserved_solution} illustrates the concepts introduced above.
+The figure shows the global minimizer $y^*$, the $\epsilon$-reserved solution $y_\epsilon$ and the set $Y_\epsilon$ of all admissible points which are not worse than the $\epsilon$-reserved solution.
 
@@
 	\center{\includegraphics[width=0.6\linewidth]{Fig1.pdf}}
-	\caption{ {\color{red} An $\varepsilon$-reserved solution to a two-dimensional problem. } }
+	\caption{ An $\varepsilon$-reserved solution to a two-dimensional problem }
 	\label{fig:eps_reserved_solution}
@@
 This situation is typical for the optimal design problems when some conditions for the functioning of the modeled object are not met and its other characteristics appear to be undefined. 
-{\color{red} In this approach, the violation of any constraint at the point $y$ terminates the computation of the values of other functions at this point. }
+In this approach, the violation of any constraint at the point $y$ terminates the computation of the values of other functions at this point.
 
@@
 	\item the global minimizer is known in advance and can be located at the boundary of the feasible domain;
-	\item {\color{red} the number of constraints active at the global minimizer can be predefined (an inequality constraint is called active at some point if at this point the inequality turns into equality)}.
+	\item the number of constraints active at the global minimizer can be predefined (an inequality constraint is called active at some point if at this point the inequality turns into equality).
 \end{itemize}

### 2020-05-27 2020/PAVT_Nested_Full/pavt_paper.tex

@@
 %
-\title{Combining of local and global search in the parallel nested optimization scheme\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
+\title{Combining Local and Global Search in a Parallel Nested Optimization Scheme\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
 }
 %
-\titlerunning{Combining of local and global search}  % abbreviated title (for running head)
+\titlerunning{Combining local and global search}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 
-The present paper considers global optimization problems and the numerical methods of solving them.
-An assumption is made regarding the dependence of the objective function on its parameters: it is multiextremal with respect to selected variables only, and the dependence on the other variables has a local character.
-Problems of this type may arise in identifying the unknown parameters of the mathematical models, based on experimental results.
+This paper considers global optimization problems and numerical methods to solve them.
+The following assumption is made regarding the dependence of the objective function on its parameters: it is multiextremal concerning selected variables only, and the dependence on the other variables has a local character.
+Such problems may arise when identifying unknown parameters of the mathematical models based on experimental results.
 A parallel computational scheme, which accounts for this feature, is being proposed.
-This novel scheme is based on the idea of nested (recursive) optimization, when at the upper recursion level the optimization with respect to the significant variables is performed (using the global optimization method), and at the lower one local optimization problems are solved (using the local method).
-The generated local subproblems will not affect each other, and solving these problems can be performed in parallel.
-Numerical experiments have been conducted on several hundred test problems, confirming the efficiency of the proposed scheme of parallel computations.
+This novel scheme is based on the idea of nested (recursive) optimization, which suggests optimizing the significant variables at the upper recursion level (using the global optimization method) and solving the local optimization problems at the lower level (using the local method).
+The generated local subproblems will not affect each other and can be solved in parallel.
+The efficiency of the proposed parallel computation scheme has been confirmed by the numerical experiments conducted on several hundreds of test problems.
 
@@
 Global optimization problems arise frequently in various fields of modern science and technology.
-For example, problems of finding the configurations of various chemical compounds corresponding to the minimum interaction energy are relevant \cite{Posypkin2014}.
+For example, the problems of finding the configurations of various chemical compounds corresponding to the minimum interaction energy are of current interest \cite{Posypkin2014}.
 In particular,  such problems arise
@@
 Another example is the problem of finding optimal control parameters for complex processes \cite{Kalyulin2017,Modorskii2016}.
-Global optimization problems arise in many other applications as well (see the review \cite{Pinter2006}).
+Many other applications also generate global optimization problems (see the review \cite{Pinter2006}).
 
-Application of global optimization methods for identifying the parameters of the mathematical models based on the data of experiments has become a classical practice.
-Problems of this type require a search for the values of the unknown parameters of the model, from which computational results according to the model are close to the results obtained experimentally.
+The application of global optimization methods for identifying the parameters of the mathematical models based on the data of experiments has become a conventional practice.
+These problems require such a search for values of the unknown model parameters, which makes the computational results based on the model close to the ones obtained experimentally.
 
-The number of parameters which are required to be identified in such a way, may reach tens and hundreds in complex mathematical models \cite{Akhmadullina2017,Nurislamova2016}.
-In problems of such dimensionality, regular search methods for global solution cannot be applied because of the extremely large computational costs for covering the search domain by the trial points.
-This remains true even with the use of efficient algorithms (for example, \cite{Evtushenko2009,Jones2009,Paulavicius2011}) constructing essentially nonuniform coverages.
+The number of parameters which are required to be identified in such a way may reach tens and hundreds in complex mathematical models \cite{Akhmadullina2017,Nurislamova2016}.
+In the problems of such dimensionality, regular search methods for a global solution cannot be applied because of the extremely large computational costs for covering the search domain by the trial points.
+This remains true even with the use of efficient algorithms (for example, \cite{Evtushenko2009,Jones2009,Paulavicius2011}) constructing essentially nonuniform coverage.
 
-However, in the problems of model identification, the parameters, which will affect the result most significantly, can often be selected.
+However, in the problems of model identification, it is usually possible to select the parameters, which will affect the result most significantly.
 As a rule, the number of such parameters is small.
-The rest of the parameters either cause an insufficient effect (in this case, multiextremality can be neglected) or dependence on the second group of parameters has a local character.
+The rest of the parameters either cause an insufficient effect (in this case, multiextremality can be neglected) or dependence on them is local.
 
-So far, the development of parallel algorithms for minimizing the essentially multidimensional functions (hundreds of variables), which accounts for the dependence of the objective function on different groups of parameters, is relevant.
-Specifically, the objective function is assumed to be multiextremal with respect to some of the variables.
-The rest of the variables have a local effect, i.e., the objective function is unimodal with respect to them.
-In this case, solving the problem can be organized according to the nested (recursive) optimization scheme.
-Solving the multiextremal subproblem, which requires the use of complex global search algorithms, will be performed at the upper level of recursion.
-The unimodal subproblems (each corresponding to a fixed set of values of the first part of parameters) will be solved at the lower level.
-For solving these subproblems, efficient methods of local optimization, linear programming, or linear algebra can be applied (subject to the character of the local subproblems).
+Currently, the development of parallel algorithms which would minimize the essentially multidimensional functions (hundreds of variables) and account for the dependence of the objective function on different groups of parameters is relevant.
+Specifically, the objective function is assumed to be multiextremal for some of the variables.
+The rest of the variables have a local effect, i.e., the objective function is unimodal for them.
+In this case, the problem-solving can be organized according to the nested (recursive) optimization scheme.
+The multiextremal subproblem, which requires the use of complex global search algorithms, will be soleved at the upper level of the recursion.
+The unimodal subproblems (each corresponding to a fixed set of values of the first part of the parameters) will be solved at the lower level.
+To solve these subproblems, efficient methods of local optimization, linear programming or linear algebra can be applied (subject to the character of the local subproblems).
 
-The text describing the results of the present study is organized as follows.
+Below you will find a description of this paper structure.
 Section 2 describes the nested optimization scheme taking into account different properties of the objective function at different nesting levels.
-A particular example of a problem of this class is presented, along with a general scheme of the organization of parallel computations.
+A particular example of such a problem is presented, along with a general scheme of the parallel computation organization.
 Section 3 is devoted to the description of the parallel global search algorithm using space-filling curves.
 The computational scheme of the parallel algorithm is presented, and its main properties are described.
-Section 4 discusses the methods of solving the local subproblems within the framework of the nested optimization scheme.
-A brief description of the properties of the local extremum search problems and known methods of solving them is given.
+Section 4 discusses the methods to solve the local subproblems within the framework of the nested optimization scheme.
+In addition, this section briefly describes the properties pertaining to the local extremum search problems and outlines the known methods for this type of problems.
 Section 5 contains the results of a numerical experiment.
-The evaluation of the scalability of the proposed parallel computation scheme in solving a series of test problems was performed.
+The scalability of the proposed parallel computation scheme for solving a series of test problems was evaluated.
 Section 6 concludes the paper.
 
-\section{Nested optimization scheme}
+\section{Nested Optimization Scheme}
 
@@
 
-Let us assume that at any fixed set of the values $\overline{x}$ the function $f(\overline{x},y)$ is a 
-multiextremal one and satisfies Lipschitz condition
+Let us assume that at any fixed set of values $\overline{x}$ the function $f(\overline{x},y)$ is a 
+multiextremal one and satisfies the Lipschitz condition
 \[
@@
 with the constant $L$ unknown a priori.
-At the same time, at any fixed set of values $\overline{y}$ the function $f(x,\overline{y})$ is a 
-unimodal one, i.e. the function $f(x,\overline{y})$ has a single minimum point $x^*$ (depending, in 
-general, on $\overline{y}$).
+At the same time, at any fixed set of values $\overline{y}$, the function $f(x,\overline{y})$ is a 
+unimodal one, i.e. the function $f(x,\overline{y})$ has a single minimum point $x^*$ (generally depending on $\overline{y}$).
 
-Accounting for such a peculiarity of the considered problem class can reduce essentially the computational costs 
-of the optimum search. Indeed, according to known nested optimization scheme 
-\cite{Carr}, the solving of the initial problem can be reduced to the problem of finding the global 
+Accounting for such a peculiarity of the considered problem class can essentially reduce the computational costs 
+of the optimum search. Indeed, according to the known nested optimization scheme 
+\cite{Carr}, the initial problem can be reduced to the problem of finding the global 
 minimum of the function $\varphi(y)$
@@
 \end{equation}
-According to (\ref{global_problem}), the computing of a single value of the function $\varphi (y)$ 
-(let us call this process a search trial) implies the solving of a unimodal problem (\ref{local_problem}), 
-which can be performed by one of the local search methods. The efficiency and sophistication of the 
-local optimization methods allow solving the subproblems (\ref{local_problem}) with the precision 
-exceeding the one of the global optimization methods essentially. Correspondingly, the problem 
-(\ref{global_problem}) can be considered as a global optimization problem, in which the objective 
-function value can be computed with high precision but this operation is time-consuming.
+According to (\ref{global_problem}), computing a single value of the function $\varphi (y)$ 
+(let us call this process a search trial) implies unimodal problem solution (\ref{local_problem}), 
+which can be performed with one of the local search methods. The efficiency and sophistication of the 
+local optimization methods allow solving subproblems (\ref{local_problem}) with the precision essentially 
+exceeding that of the global optimization methods. Correspondingly, the problem 
+(\ref{global_problem}) can be considered a global optimization problem, where the objective 
+function can be computed with high precision but this operation is time-consuming.
 
-The following approximation problem can be considered as a particular example of a problem, in 
-which different character of dependence on different groups of parameters takes place. 
-Let us assume $m$ values $u_j = u(t_j), 1 \leq j \leq m, $ of the function $u(t)$ to have been obtained 
-in the experiment and the analytical expression for $u(t)$ is known to take the form 
+The following approximation problem can be considered a particular example of the problem, where 
+a different character of dependence on different groups of parameters takes place. 
+Let us assume $m$ values $u_j = u(t_j), 1 \leq j \leq m, $ of the function $u(t)$ have been obtained 
+in an experiment and the analytical expression for $u(t)$ is known to take the form 
 \begin{equation}\label{ex_func}
@@
 $\omega=\left(\omega_0,\omega_1,\ldots,\omega_n\right)$ are the unknown parameters defining 
-particular function $u(t)$.
+a particular function $u(t)$.
 
-Let us introduce a measure of the deviation of the function $u(t,\omega,c)$ from the experiential data as a 
+Let us introduce a measure of function $u(t,\omega,c)$ deviation from the experiential data as a 
 sum of squares 
@@
 
-The problem (\ref{ex_prob}) can be written in a recursive form 
+Problem (\ref{ex_prob}) can be written in a recursive form 
 \begin{equation}\label{outer_prob}
@@
 \end{equation}
-The nested subproblem (\ref{inner_prob}) is a classical linear least squares problem, and its solution 
+Nested subproblem (\ref{inner_prob}) is a classical linear least squares problem, and its solution 
 can be obtained by solving a system of linear algebraic equations regarding the unknown $c$, which 
-can be done e.g. by Gaussian elimination. At the same time, the external problem (\ref{outer_prob}) 
+can be done, for example, by Gaussian elimination. At the same time, external problem (\ref{outer_prob}) 
 will be, in general, a multiextremal one, and its solving will require considerable 
@@
 
-So far, to solve the problem (\ref{main_problem}) in the nested scheme (\ref{global_problem}), 
-(\ref{local_problem}) one can apply parallel global search algorithm for solving the external 
+So far, to solve problem (\ref{main_problem}) in nested scheme (\ref{global_problem}), 
+(\ref{local_problem}) one can apply a parallel global search algorithm for solving external 
 subproblems (\ref{global_problem}). At that, a set of independent subproblems (\ref{local_problem}) 
@@
 using the local methods.
-A general scheme of the organization of the computations using several processors is presented in 
+A general scheme of the computation organization using several processors is presented in 
 Fig.~\ref{MPI_tree}.
@@
 
-The processes of the parallel program will form a tree. 
-The root process will solve the problem (\ref{global_problem}) and distribute the subproblems 
-(\ref{local_problem}) among the child processes. Each subproblem is solved by a separate process, 
+The parallel program processes will form a tree. 
+The root process will solve problem (\ref{global_problem}) and distribute subproblems 
+(\ref{local_problem}) among the child processes. Each subproblem is solved by a separate process; 
 the data exchange between the processes can be organized using MPI. The data transmission between 
 the processes will be minimal: it is required to send the coordinates of points $y^1, ..., y^p$ from the 
-root to the child processes and to receive back the found minimal values 
+root to the child processes and to receive the found minimal values back
 \[
@@
 
-\section{Parallel global search algorithm}
+\section{Parallel Global Search Algorithm}
 
-An efficient parallel global optimization algorithm intended for solving the problems of the kind 
+An efficient parallel global optimization algorithm intended for solving the problems of kind 
 (\ref{global_problem}) has been developed in Lobachevsky University of Nizhni Novgorod (UNN) 
 \cite{Barkalov2018,globalizerSystem,Strongin2018}. 
-The main idea of the parallelization consists in such of organization of computations that several trials 
-are performed in parallel. This approach is featured by high efficiency (the part of the computational 
-process is parallelized, in which the major amount of computations is performed) and generality (it is 
+The main idea of the parallelization consists in such an organization of computations during which several trials 
+are performed in a parallel way. This approach is featured by a high efficiency (the part of the computational 
+process with the major amount of computations is parallelized) and generality (it is 
 applicable for a wide class of algorithms).
 
-In the developed global search algorithm, a novel method for the reduction of the dimensionality of 
+In the developed global search algorithm, a novel method to reduce the dimensionality of 
 the problems to be solved is applied. 
-The reduction of the dimensionality is based on a fundamental fact, according to which an 
-$N$-dimensional hyperparallelepiped $D$ and the interval $[0,1]$ of the real axis are the 
+The reduction of the dimensionality is based on the fundamental fact, according to which an 
+$N$-dimensional hyperparallelepiped $D$ and the interval $[0,1]$ of the real axis are  
 equal-power sets, and the interval $[0,1]$ can be mapped continuously into $D$ using the Peano 
@@
 
-Utilizing this fact, one can reduce the solving of the problem of minimization of a multidimensional 
+Utilizing this fact, one can reduce the minimization problem of a multidimensional 
 function $\varphi(y)$ to the minimization of a one-dimensional function $\varphi(y(x))$
@@
 with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation
-$ H=2 L \sqrt{N+3}$ and $y(x)$ is the Peano curve from $[0,1]$ onto $D$.
+$ H=2 L \sqrt{N+3}$ and $y(x)$ being the Peano curve from $[0,1]$ onto $D$.
 
@@
 are given in \cite{Sergeyev2013,Strongin2000}. As an illustration, two evolvents are shown in 
-Fig.~\ref{fig_evolvent}. Figure demonstrates that the precision of the evolvent is depend on the 
+Fig.~\ref{fig_evolvent}. The figure demonstrates that the precision of the evolvent depends on the 
 density level $m$ used in the construction.
@@
 \end{minipage}
-\caption{Evolvents in two dimensions with (a) $m=4$ and (b) $m=5$}
+\caption{The evolvents in two dimensions with (a) $m=4$ and (b) $m=5$}
 \label{fig_evolvent}
@@
 
-It is worth noting that many well known global optimization algorithms are based implicitly on the 
-idea of the dimensionality reduction and adaptation of the one-dimensional algorithms for solving the 
+It is worth noting that many well-known global optimization algorithms are based implicitly on the 
+idea of the dimensionality reduction and adaptation of one-dimensional algorithms for solving  
 multidimensional problems \cite{Evtushenko2013,Sergeyev2006,Zilinskas2008}.
 
-According to \cite{Strongin2018,Strongin2000}, the rules of the global search algorithm, in which 
+According to \cite{Strongin2018,Strongin2000}, the rules of the global search algorithm, where 
 $p$ trials are performed in parallel at every iteration, are as follows.
@@
 
-Suppose $n\geq 1$  iterations of the method have been carried out, in which trials were performed at 
-$k=k(n)$ points $x^i,1\leq i \leq k$. Then points $x^{k+1},\dots,x^{k+p}$  of search trials of the next 
+Supposing $n\geq 1$ iterations of the method have been carried out, where the trials were performed at 
+$k=k(n)$ points $x^i,1\leq i \leq k$, then the points $x^{k+1},\dots,x^{k+p}$  of the search trials at the next 
 iteration are determined according to the following rules.
 
-Rule 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the
-lower indices in ascending order of coordinate values, i.e.
+Rule 1. Renumber the points $x^1,...,x^k$ of the preceding trials in lowercase in ascending order of the coordinate values, i.e.
 \[
@@
 \]
-and juxtapose to them the values $z_i=\varphi(y(x_i)), \; 1 \leq i \leq k$, computed at these points.
+and juxtapose them with the values $z_i=\varphi(y(x_i)), \; 1 \leq i \leq k$, computed at these points.
 
@@
 \end{equation}
-where $\Delta_i=(x_i-x_{i-1})^{1/N}$. The value $r > 1$ is parameter of the algorithm. An 
-appropriate selection
-of $r$ allows to consider the product $r \mu$ as an estimate
+where $\Delta_i=(x_i-x_{i-1})^{1/N}$. The value $r > 1$ is an algorithm parameter. An 
+appropriate selection of $r$ allows considering the product $r \mu$ as an estimate
 of the Lipschitz constant $L$ of the objective function.
@@
 Rule 5. Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), \; 1\leq j\leq p$, calculated 
-using the formulae
+using the formula
 \begin{equation}\label{Rule_X}
@@
 
-The algorithm terminates if the condition $\Delta_{t_j}<\epsilon$ is satisfied at least for one number 
+The algorithm terminates if the condition $\Delta_{t_j}<\epsilon$ is performed at least for one number 
 $t_j, 1 \leq j \leq p$ ; $\epsilon>0$ is the predefined accuracy.
@@
 point location in these intervals. Inequalities (\ref{Rule_Max}) arrange intervals according to their 
-characteristics, and trials are carried out in parallel in the first $p$ intervals with the largest 
+characteristics, and the trials are carried out in parallel in the first $p$ intervals with the largest 
 probabilities.
@@
 %Замечание по поводу теоретических свойств и сравнения алгоритмов
-The results of the experimental comparison  (see, for example, \cite{Sovrasov2019}) show that global search algorithm is superior to many known methods of similar purpose, including both deterministic and heuristic  algorithms.
+The results of the experimental comparison  (see, for example, \cite{Sovrasov2019}) show that the global search algorithm is superior to many known methods of similar purpose, including both deterministic and heuristic  algorithms.
 
-\section{Local search algorithms}
+\section{Local Search Algorithms}
 
-In the present section, let us consider the issues related to the solving the subproblem 
+In this section, let us consider the issues related to solving subproblem 
 (\ref{local_problem}), which (in the notation $f(x) = f(x,\overline{y})$ for the fixed value 
-$\overline{y} \in D$) corresponds the problem of finding a local extremum
+$\overline{y} \in D$) corresponds to the problem of finding a local extremum
 
@@
 
-To date, a huge number of various local optimization methods for the problems of the kind (\ref{lp}) 
-have been developed. Among these ones, there are, for example, the gradient methods, Newton and 
-quasi-Newton methods, the methods of conjugate directions, etc. The majority of these ones utilize 
-the principle of local descent, when the method passes to the points with smaller values of the 
-objective function at every step progressively. Almost all these methods can be represented in the 
+To date, a huge number of various local optimization methods for the problems of kind (\ref{lp}) 
+have been developed. They include such methods as the gradient ones, Newton, and 
+quasi-Newton methods, the methods of conjugate directions, etc. The majority of them utilizes 
+the principle of local descent when the method progressively passes the points with smaller values of the 
+objective function at every step. Almost all these methods can be represented in a 
 form of the iterative relation 
@@
 where $x^k$ are the points of main trials consisting in computing a set of some local characteristics of 
-the objective function $I^k=I(x^k)$ at the point $x^k$, where $d^k$ are the directions of shift from 
-the points $x^k$ computed on the base of the results of main trials and $s^k$ are the coefficients 
-defining the magnitudes of the shifts along to the directions chosen.
+the objective function $I^k=I(x^k)$ at the point $x^k$, where $d^k$ are the shift directions from 
+the points $x^k$ computed on the base of the main trial results, and $s^k$ are the coefficients 
+defining the magnitudes of the shifts along the directions chosen.
 
-In order to define the shift magnitudes $s^k$ along the directions $d^k$, the methods can perform 
-auxiliary (working) steps. This results in additional measuring of the local characteristics of the 
-objective function along the direction $d^k$. The transitions from the points $x^k$ to the points 
-$x^{k+1}$ is performed in such a way that an essential decrease of the value of the function $f^k = 
-f( x^k )$ is provided as a result of the step.
+To define the shift magnitudes $s^k$ along the directions $d^k$, the methods can include some  
+auxiliary (working) steps. This results in additional measuring of the objective function local characteristics  
+along the direction $d^k$. The transitions from the points $x^k$ to the points 
+$x^{k+1}$ is performed in such a way that the function $f^k = f( x^k )$ value essentially decreases as a result of this step.
 
 The set of the computed local characteristics $I^k=I(x^k)$ may include: the function values $f^k = 
-f( x^k )$, the gradient $\nabla f^k = \nabla f(x^k)$, the matrix of the second derivatives 
-(Hessian) $\Gamma_k=\Gamma^f(x^k)$. Particular set of the characteristics to be measured depends 
+f( x^k )$, the gradient $\nabla f^k = \nabla f(x^k)$, and the matrix of the second derivatives 
+(Hessian) $\Gamma_k=\Gamma^f(x^k)$. A particular set of the characteristics to be measured depends 
 on the properties of the problem being solved as well as on the optimization method chosen.
@@
 In the local optimization problems arising in the applications, the a priori information on the objective 
-functions is usually quite limited. For example, a local character of the dependence of the function on 
-the parameters may be suggested only whereas the gradient values (and, moreover, Hessian) are
-unknown. In this case, the direct search methods are applied, which don't utilize any suggestions on
-the smoothness of the objective function. When searching for the minimum, these methods measure 
-the function values only. The rules of the placement of the iteration points in these ones is based on 
+functions is usually quite limited. For example, a local character of the function dependence on 
+the parameters may only be suggested where the gradient values (and Hessian) are
+unknown. In this case, the applied methods are the direct search ones, which do not utilize any suggestions on
+the objective function smoothness. When searching for the minimum, these methods measure 
+the function values only. The rules of the placement of the iteration points in these ones are based on 
 some heuristic logical schemes.
@@
 Hooke--Jeeves method \cite{HookJeeves} and Nelder--Mead method \cite{NelderMead} belong to 
-the popular methods of the direct search. In spite of apparent simplicity and theoretical 
-non-substantiation of the direct search methods, these ones are well established in the practical 
-computations. This can be explained as follows. Many methods of smooth optimization are very 
-sensitive to the computational errors in the function values transforming theoretically smooth function 
-into the non-smooth one actually. Because of this, in the practical computations these ones loose the 
-positive properties, which are promised by theory often. The use of the direct search methods allows 
-achieving better results in these conditions.
+the popular methods of the direct search. Despite their apparent simplicity and theoretical 
+non-substantiation of the direct search methods, these two methods are well established in the practical 
+computations. This can be explained as follows. Actually, many smooth optimization methods are very 
+sensitive to computational errors in the function values transforming the theoretically smooth function 
+into the non-smooth one.
+Because of this, in the practical computations, these methods lose positive properties, which are often promised by theory.
+In these conditions, direct search methods allow achieving better results.
 
-\section{Numerical experiments}
+\section{Numerical Experiments}
 
@@
 Globalizer software system developed in UNN \cite{globalizerSystem,Sysoyev2017}.  
-The numerical experiments were carried out using Lomonosov supercomputer (Lomonosov Moscow 
+The numerical experiments were carried out using the Lomonosov supercomputer (Lomonosov Moscow 
 State University). Each supercomputer node included two quad-core processors Intel Xeon X5570 
-2.93GHz and 12 Gb RAM. When building the Globalizer system for running on the Lomonosov 
-supercomputer GCC 5.5.0 compiler and Intel MPI 2017 were used.
+2.93GHz and 12 Gb RAM. To build the Globalizer system for running on the Lomonosov 
+supercomputer, we used the GCC 5.5.0 compiler and Intel MPI 2017.
 
-In order to simulate the applied problems, in which the locally-affecting parameters can be selected, 
-the test function of the kind 
+To simulate the applied problems, where the locally-affecting parameters can be selected, 
+a test function of the kind 
 \begin{equation}\label{test_problem}
@@
 \end{equation}
-were used when conducting the experiments. Here $G(y)$ is a multiextremal function of the 
+was used when conducting the experiments. Here, $G(y)$ is a multiextremal function of the 
 dimensionality $N=2$ and $R(x)$ is a unimodal function of the dimensionality $N \gg 2$.
@@
 \end{eqnarray}
-and the coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$ are the random numbers distributed uniformly 
+and the coefficients $A_{ij}, B_{ij}, C_{ij}, D_{ij}$ are the random numbers uniformly distributed  
 in the interval $[-1,1]$.
-This class of functions is often used for testing the global optimization algorithms. 
+This class of functions is often used for testing global optimization algorithms. 
 As the local part of the problem, the modified Rosenbrock function 
@@
 \]
-was used, in which the extremum point was shifted in the search domain randomly using the linear 
+was used, where the extremum point was randomly shifted in the search domain using the linear 
 coordinate transform.
 
-As an example, a contour plot of a pair of functions  $G(y)$ and $R(x)$ is presented in 
-Fig.~\ref{fig_level}(a),(b). These functions are the complex ones for corresponding global and local 
-optimization methods since the function $G(y)$ is essentially multiextremal one, and the function 
-$R(x)$ has a ravine structure manifested clearly.
+As an example, contour plots of a pair of functions  $G(y)$ and $R(x)$ are presented in 
+Fig.~\ref{fig_level}. These functions are the complex ones for the corresponding global and local 
+optimization methods since the function $G(y)$ is essentially multiextremal, and the function 
+$R(x)$ has a ravine clearly manifested structure.
 
@@
 In Fig.~\ref{fig_level}, the trial points performed by the global search algorithm and Hooke--Jeeves 
-method when solving corresponding problems are marked also. 
-The positions of the trial points demonstrate clearly the multiextremality of the function $G(y)$ (the 
+method when solving the corresponding problems are also marked. 
+The trial point positions clearly demonstrate the function $G(y)$ multiextremality (the 
 trial points form a nonuniform coverage of the search domain) as well as the ravine character 
@@
 
-The problem was considered to be solved if in the course of minimization of the function $G(y)$, the 
-global search algorithm generates the next trial point $y^k$ in the $\epsilon$-nearness of the global 
+The problem was considered to be solved if in the course of minimizing the function $G(y)$, the 
+global search algorithm generates the next trial point $y^k$ in the $\epsilon$-neighbourhood of the global 
 minimum $y^*$, i.e. $\left\|y^* - y^k\right\|\leq\epsilon$, where $\epsilon = 10^{-2}$. 
@@
 In the global search algorithm, the parameter $r=4.5$ from (\ref{Rule_R}) and the evolvent building 
-parameter $m=10$ were used; the local Hooke--Jeeves method employed has no additional parameters.
+parameter $m=10$ were used; the employed local Hooke--Jeeves method had no additional parameters.
 
-In the first series of experiments, 100 problems of the kind (\ref{test_problem}) were solved, in which 
-the dimensionality of the local subproblems was $N=50$. The problems were solved using 1, 4, and 8 
-cluster nodes, in which from 2 up to 32 MPI-processes were employed. According to the 
-parallelization scheme presented in Fig.~\ref{MPI_tree}, the solving of the global optimization 
-problem at the upper nesting level was performed in a single (master) process, which initiated the 
-parallel solving of the local problems of the lower level (from 1 up to 31 problems, respectively). So 
-far, the minimum possible number of processes employed was equals to 2, one per each nesting level.
+In the first series of experiments, 100 problems of kind (\ref{test_problem}) were solved.  Their local subproblems dimensionality was $N=50$. The problems were solved using 1, 4, and 8 
+cluster nodes, which employed from 2 to 32 MPI-processes. According to the 
+parallelization scheme presented in Fig.~\ref{MPI_tree}, at the upper nesting level the global optimization problem was solved in a single (master) process, which initiated parallel solving of the lower-level local problems (from 1 up to 31 problems, respectively).
+So far, the minimum possible number of the employed processes equalled 2, one per each nesting level.
 
-Table \ref{tab1} presents the averaged number of iterations $K_{av}$ performed by the global search 
-algorithm when minimizing the function $G(y)$, the averaged time of solving the whole problem 
-$T_{av}$, and the speed-up in time $S$ subject to the number of employed MPI-processes $p$.
+Table \ref{tab1} presents the average number of iterations $K_{av}$ performed by the global search 
+algorithm when minimizing the function $G(y)$, the average time of solving the whole problem 
+$T_{av}$, and the speed-up $S$ subject to the number of employed MPI processes $p$.
 
@@
 
-In the second series of experiments, 10 problems of the kind (\ref{test_problem}) were solved, in 
-which the dimensionality of the local subproblems was equal $N=100$. The running parameters were 
+In the second series of experiments, 10 problems of kind (\ref{test_problem}) were solved, in 
+which the dimensionality of the local subproblems was equal to $N=100$. The running parameters were 
 the same as in the previous series of experiments. The results of solving are presented in Table 
@@
 \centering
-\caption{Results of solving a series of problems of dimensionality $N=102$}\label{tab2}
+\caption{The results of solving a problems series with dimensionality $N=102$}\label{tab2}
 \begin{tabular}{cccc}
@@
 This class of functions is often used for investigating global optimization algorithms \cite{Barkalov2018,Paulavicius2014,Sergeyev2015}.
-The dimensionality of the local subproblems was equal $N=100$.
-The problems were solved at 1, 8, 16 and 32 cluster nodes, on which from 4 to 128 MPI processes were used.
+The dimensionality of the local subproblems was equal to $N=100$.
+The problems were solved at 1, 8, 16, and 32 cluster nodes, on which from 4 to 128 MPI processes were used.
 The running parameters were the same as in the previous series of experiments.
-The results of solving are presented in Table \ref{tab3}, where speedup is calculated with respect to  running on a single node using 4 MPI processes.
+The results of solving are presented in Table \ref{tab3}, where speed-up is calculated for running on a single node using 4 MPI processes.
 
@@
 \centering
-\caption{Results of solving a series of problems of dimensionality $N=103$}\label{tab3}
+\caption{The results of solving a problems series with dimensionality $N=103$}\label{tab3}
 \begin{tabular}{cccc}
@@
 
-Results of experiments performed on the series of optimization problems demonstrate that the scheme of parallel computations proposed in the present work has good potential for parallelization.
-Specifically, when solving a series of 100 problems with a total number of variables equal to 52 (from which the two first variables caused a global effect on the objective function and fifty affected the one locally), a speedup of 24.6 was obtained when using 32 processes.
-When solving  problems with more complex local subproblems (with the number of the local variables N=100), the speedup increased up to 35 when using 128 processes.
-In the last case the speedup was calculated with respect to the use of 4 processes on one node.
-The developed scheme of parallel computations can be applied in identifying complex mathematical models, which are featured by large numbers (tens and hundreds) of unknown parameters.
+The results of the experiments performed on a series of optimization problems demonstrate that the parallel computation scheme proposed in this work has a good potential for the parallelization.
+Specifically, when solving a series of 100 problems with a total number of variables equal to 52 (from which the first two variables caused a global effect on the objective function and the other fifty affected the one locally), the obtained speed-up was 24.6 when using 32 processes.
+When solving problems with more complex local subproblems (with the local variables number $N=100$), the speed-up increased up to 35 using 128 processes.
+In the last case, the speed-up was calculated subject to using 4 processes on one node.
+The developed parallel computation scheme can be applied in identifying complex mathematical models, which feature a large number (tens and hundreds) of unknown parameters.
+
 %

### 2020-05-27 2020/RSD_GPU/RSD2020.tex
Исправлена ошибка в таблице с шестимерными задачами.
Add calculation time

@@
 Therefore, in order to simulate the computational complexity inherent in applied optimization problems \cite{Modorskii2016}, the calculation of the objective function in all experiments was complicated by additional calculations that do not change the form of the function and the location of its minima (summing the segment of the Taylor series).
+In the experiments carried out the average time for calculating the function values was 0.01 seconds.
 
@@
                                \noalign{\smallskip} \hline \noalign{\smallskip}                               
-                               $p=64$  &    11,3	 & & 6,8	  \\
-                               $p=256$ &    39,0	 & & 31,5  \\
-															 $p=1024$&    128,4	 & & 83,5  \\
+                               $p=64$  &    30,8	 & & 1,9	  \\
+                               $p=256$ &    92,7	 & & 1,5  \\
+							   $p=1024$&    597,0	 & & 2,5  \\
                                \noalign{\smallskip}\hline
@@
 
-The last series of experiments has been carried out on 100 six-dimensional problems from the GKLS Simple class. 
+The last series of experiments has been carried out on 20 six-dimensional problems from the GKLS Simple class. 
 Table \ref{table:GKLS6_average_MMAS_GPU} shows the speedup of the algorithm combining multiple evolvents and the adaptive scheme when solving the problems on 3 cluster nodes (using 2 GPUs per node, 6144 GPU threads in all) compared to the adaptive scheme on CPU using 4 threads.

### 2020-05-27 2020/PAVT_Sovrasov/slides/slides.tex
PAVT: update slides

@@
           \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
-          D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
+          D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N},\:g_j(y   )\leqslant 0, 1\leqslant j\leqslant m\}
         \end{array}
       \end{displaymath}
-      \(\varphi(y)\) -- многоэкстремальная функция, удовлятворяющая условию Липшица:
+      \(\varphi(y),\:g_j(y)\) -- многоэкстремальные функции, удовлятворяющие условию Липшица:
       \begin{displaymath}
-        |\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
+        |f(y_1)-f(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
       \end{displaymath}
@@
   \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Постановка задачи}
+  Далее будем интересоваться решением серии из \(q\) задач глобальной оптимизации с нелинейными ограничениями:
+  \begin{displaymath}
+    \min\left\{\varphi_1(y), y\in D_1 \right\}, \min\left\{\varphi_2(y), y\in D_2\right\},..., \min\left\{\varphi_q(y), y\in D_q\right\}.
+  \end{displaymath}
+  Подобные серии задач могут возникнуть, например, в следующих случаях:
+  \begin{itemize}
+    \item задача глобальной оптимизации с дискретным параметром;
+    \item  решение задачи многокритериальной оптимизации методом свертки критериев.
+  \end{itemize}
+\textbf{Возможные методы решения}:
+  \begin{itemize}
+    \item решать каждую задачу независимо;
+    \item разработать метод оптимизации, который решает все задачи из множества в совокупности, в каждый
+    момент времени приоретизируюя одну из них.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \begin{center}
+  \frametitle{Редукция размерноти}
+  Кривая Пеано \(y(x)\) позволяет уменьшить размерность многомерного пространства до 1:
+  \begin{gather}
+    \lbrace y\in \mathbb{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\} \nonumber \\
+    \min\{\varphi(y): y\in D\}=\min\{\varphi(y(x)): x\in [0,1]\} \nonumber
+  \end{gather}
+  \(y(x)\) является негладкой кривой, отображающей отрезок \([0,1]\) на гиперкуб \(D\).
+  \begin{figure}[ht]
+    \vspace*{-0.5cm}
+    \subfloat{{\includegraphics[width=.35\textwidth]{peano3d.png} }}
+    \subfloat{{\includegraphics[width=.35\textwidth]{peano2d.png} }}
+  \end{figure}
+\end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Алгоритм глобальной оптимизации}
+  Метод оптимизации генерирует последовательность точек \(\{x_k:x_k\in[a,b]\}\) и состоит в выполнении следующих шагов:
+  \begin{enumerate}
+    \setlength{\itemindent}{.1in}
+    \item[Шаг 1.] Упорядочить поисковую информацию (одномерные точки) по возрастанию.
+    \item[Шаг 2.] Для каждого интервала \((x_{i-1}, x_i)\) вычислить величину \(R(i)\), называемую характеристикой.
+    \item[Шаг 3.] Выбрать интервал \((x_{t-1}, x_{t})\) с наибольшей характеристикой и
+    провести испытание (вычислить органичения и целевую функцию) в точке \(x^{k+1}\), выбранной с помощью решающего правила \(d\):
+    \begin{displaymath}
+      x^{k+1}=d(t)\in (x_{t-1}, x_{t})
+    \end{displaymath}
+    \item[Шаг 4.] Если \(x_{t}-x_{t-1}<\varepsilon\), остановить метод.
+  \end{enumerate}
+  \textit{\footnotesize	{Детальное описание метода: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Алгоритм, решающий множество задач}
+  Создать \(q\) копий характеристического АГО.
+  использовать \(q\) синхронно работающих копий ИАГП с тем лишь отличием, что на шаге 6 при выборе
+  интервала с наилучшей характеристикой, выбор будет осуществляться из всех интервалов, которые
+  породили на данный момент \(q\) копий ИАГП. Если наибольшая характеристика соответствует
+  задаче \(i\), то выполняется шаг 7 в копии метода с номером \(i\), а остальные копии метода простаивают.
+
 \end{frame}

### 2020-05-27 2020/RSD_GPU/RSD2020.tex
12 pages again

@@
 
-Thus, calculating the optimization criterion can be implemented on GPU, and the role of the optimization algorithm (running on CPU) consists in the effective selection of points for conducting parallel trials.
-This scheme of working with the accelerator is fully consistent with the work of the parallel global search algorithm developed at the Lobachevsky State University of Nizhni Novgorod and presented in a series of publications \cite{Barkalov2014,Barkalov2020,Barkalov2019,Grishagin2016,Grishagin2015,Grishagin2018,globalizerSystem}.
+Thus, calculating the optimization criterion can be implemented on GPU, and the role of the optimization algorithm (running on CPU) consists in the selection of points for conducting parallel trials.
+This scheme of working with the accelerator is fully consistent with the work of the parallel global search algorithm developed at the Lobachevsky State University of Nizhni Novgorod and presented in a series of papers \cite{Barkalov2014,Barkalov2020,Barkalov2019,Grishagin2016,Grishagin2015,Grishagin2018,globalizerSystem}.
 
@@
 
-In this paper, we will use an approach based on the idea of reducing dimensionality using the Peano space-filling curve \(y(x)\), which continuously and unambiguously maps a segment of the real axis \([0,1]\) onto an \(n\)-dimensional cube
+In this work we will use an approach based on the idea of reducing dimensionality using the Peano space-filling curve \(y(x)\), which continuously and unambiguously maps a segment of the real axis \([0,1]\) onto an \(n\)-dimensional cube
 \begin{equation}
@@
 
-The questions of numerical construction of approximations to Piano curve (evolvents) and the corresponding theory are discussed in detail in \cite{Sergeyev2013,Strongin2000}.
+The questions of numerical construction of approximations to Peano curve (\textit{evolvents}) and the corresponding theory are discussed in detail in \cite{Sergeyev2013,Strongin2000}.
 Using evolvents \(y(x)\) reduces the multidimensional problem (\ref{task}) to a one-dimensional problem
@@
 Data transfers from CPU to GPU will be minimal: one need only transfer the coordinates of the trial points to GPU, and get back the function values at these points.
-Functions that determine the processing of trial results in accordance with the algorithm, and require work with large amounts of accumulated search information, can be effectively implemented on CPU.
+Functions that determine the processing of trial results in accordance with the algorithm, and require work with large amounts of accumulated search information, can be efficiently implemented on CPU.
 
@@
 
-   \begin{figure}[h!]
+   \begin{figure}[t]
     \centering
-		\includegraphics[width=0.9\textwidth]{GPU_Algorithm.pdf}
+		\includegraphics[width=0.75\textwidth]{GPU_Algorithm.pdf}
 		\caption{Scheme of information exchanges in the GPU algorithm}\label{fig:GPU_Algorithm}
@@
 
-   \begin{figure}[h!]
+   \begin{figure}[t]
     \centering
-		\includegraphics[width=0.7\textwidth]{Parallel_Computing.pdf}
+		\includegraphics[width=0.65\textwidth]{Parallel_Computing.pdf}
 		\caption{Diagram of parallel computing on a cluster}\label{fig:Parallel_Computing}
@@
 Therefore, in order to simulate the computational complexity inherent in applied optimization problems \cite{Modorskii2016}, the calculation of the objective function in all experiments was complicated by additional calculations that do not change the form of the function and the location of its minima (summing the segment of the Taylor series).
-In the experiments carried out the average time for calculating the function values was 0.01 seconds.
+In the experiments carried out the average time for calculating the function values was 0.01 seconds, which exceeds the latency of the network or the data transfer time to the GPU.
 
@@
 
-\begin{table}
+\begin{table}[t]
 	\caption{Average number of iterations $k_{av}$}\label{table:average_iters}
@@
 
-\begin{table}
+\begin{table}[t]
                 \caption{Speedup on CPU}
@@
 
-\begin{table}
+\begin{table}[t]
                 \caption{Speedup on one GPU}
@@
 
-\begin{table}
+\begin{table}[t]
                 \caption{Speedup on two GPUs}
@@
 
-\begin{table}
+\begin{table}[t]
                 \caption{Speedup on six GPUs}

### 2020-05-27 2020/PAVT_Sovrasov/slides/slides.tex
PAVT: pre-final version of slides

@@
 \inputencoding{utf8}
-\usepackage[russian]{babel}
 \usepackage{listings}
@@
           \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
-          D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N},\:g_j(y   )\leqslant 0, 1\leqslant j\leqslant m\}
+          D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, \\ 1\leq{i}\leq{N},\:g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}
         \end{array}
@@
 \begin{frame}
+  \frametitle{Редукция размерноти}
   \begin{center}
-  \frametitle{Редукция размерноти}
   Кривая Пеано \(y(x)\) позволяет уменьшить размерность многомерного пространства до 1:
@@
   \begin{figure}[ht]
-    \vspace*{-0.5cm}
-    \subfloat{{\includegraphics[width=.35\textwidth]{peano3d.png} }}
-    \subfloat{{\includegraphics[width=.35\textwidth]{peano2d.png} }}
+    \includegraphics[width=.35\textwidth]{peano3d.png}
+    \includegraphics[width=.35\textwidth]{peano2d.png}
   \end{figure}
@@
   \frametitle{Алгоритм, решающий множество задач}
-  Создать \(q\) копий характеристического АГО.
-  использовать \(q\) синхронно работающих копий ИАГП с тем лишь отличием, что на шаге 6 при выборе
+  Модификация метода для решения множества задач:
+  \begin{itemize}
+    \item Создать \(q\) копий характеристического АГО.
+    \item Использовать \(q\) синхронно работающих копий АГО с тем лишь отличием, что на шаге 3 при выборе
     интервала с наилучшей характеристикой, выбор будет осуществляться из всех интервалов, которые
-  породили на данный момент \(q\) копий ИАГП. Если наибольшая характеристика соответствует
-  задаче \(i\), то выполняется шаг 7 в копии метода с номером \(i\), а остальные копии метода простаивают.
+    породили на данный момент \(q\) копий АГО.\
+    \item Если наибольшая характеристика соответствует
+    задаче \(i\), то выполняется шаг 3 в копии метода с номером \(i\), а остальные копии метода простаивают.
+    \item Для реализации параллельности в предыдущем пункте можно выбрать \(p\) характеристик и провести
+    испытания в \(p\) задачах одновременно.
+  \end{itemize}
+  Чтобы приведённая схема работала, требуется \textbf{сравнимость характеристик интервалов в разных задачах}.
+\end{frame}
 
+\begin{frame}
+  \frametitle{Индексная схема учёта ограничений}
+  Вместо исходной задачи рассмотрим следующую задачу без функциональных ограничений:
+  \begin{displaymath}
+    \begin{array}{lr}
+      \psi (x^{*})=\min_{x\in [0;1]}\psi (x), \\
+      \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu }&\nu <M\\(g_{M}(x)-g_{M}^{*})/H_{M}&\nu =M\end{cases}}
+    \end{array}
+  \end{displaymath}
+  где при \(\nu=m+1\;g_\nu(x)=\varphi(x)\).
+
+  При использовании индексной схемы характеристики определяются следующим образом:
+  \begin{displaymath}
+    R(i)={\begin{cases}\Delta _{i}+{\frac {(z_{i}-z_{i-1})^{2}}{(r_{\nu }\mu _{\nu })^{2}\Delta _{i}}}-2{\frac {z_{i}+z_{i-1}-2z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu (x_{i})=\nu (x_{i-1})\\2\Delta _{i}-4{\frac {z_{i-1}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu (x_{i-1})>\nu (x_{i})\\2\Delta _{i}-4{\frac {z_{i}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu (x_{i})>\nu (x_{i-1})\end{cases}}
+  \end{displaymath}
+  где \(z_i=\psi(x_i)\). Можно показать, что при использовании таких характеристик метод будет выбирать интервалы из всех \(q\) задач, тем самым решая их все, а не только некоторое подмножество задач.
+\end{frame}
+
+\begin{frame}
+  \frametitle{Пример решения многокритериальной задачи}
+  Рассматриваемая задача:
+  \begin{displaymath}
+    \begin{array}{l}
+        Minimize \left \{
+        \begin{array}{l}
+          f_1(y) = 4 y_1^2 + 4 y_2^2 \\
+          f_2(y) = (y_1-5)^2 + (y_2-5)^2 \\
+        \end{array}
+        \right .
+        y_1\in [-1;2],y_2\in [-2;1]
+        \\s.t.
+        \\
+        \left \{
+        \begin{array}{l}
+          g_1(y) = (y_1 - 5)^2 + y_2^2 - 25 \leqslant 0 \\
+          g_2(y) = -(y_1 - 8)^2 - (y_2 + 3)^2 + 7.7 \leqslant 0\\
+        \end{array}
+        \right .
+    \end{array}
+  \end{displaymath}
+  После использования свёртки Гермейера для скаляризации задача примет вид:
+  \begin{displaymath}
+    \varphi(y,\lambda_1,\lambda_2)=\max\{\lambda_1 f_1(y), \lambda_2 f_2(y):\lambda_1,\lambda_2\in[0,1],\: \lambda_1+\lambda_2=1\}
+\end{displaymath}
+Для численного построения множества Парето выберем
+100 наборов коэффициентов \((\lambda_1,\lambda_2)\) таких, что
+\(\lambda_1^i=i h,\: \lambda_2^i=1-\lambda_1^i,\: h=10^{-2},i=\overline{1, 100}\).
+\end{frame}
+
+\begin{frame}
+  \frametitle{Пример решения многокритериальной задачи}
+  \vspace*{-1.0cm}
+  \begin{displaymath}
+    \begin{array}{cr}\\
+      SP(S)=\sqrt{\frac{1}{|S|-1} \sum_{i=1}^{|S|} (\overline{d}-d_i)},\:\overline{d}=mean\{d_i\}, \\
+      d_i=\min_{s_i,s_j\in S:s_i\ne s_j}||F(s_i)-F(s_j)||_1,\: F=(f_1,f_2)
+    \end{array}
+  \end{displaymath}
+
+  \begin{figure}[ht]
+      \centering
+      \vspace*{-0.5cm}
+      \subfloat[Раздельное решение задач, \(SP_{single}=0.984\)]{
+      \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{single_mco.pdf}}}
+      \subfloat[Решение множества задач, \(SP_{multi}=0.749\)]{
+      \label{fig:mco_pareto_2} {\includegraphics[width=.5\textwidth]{multi_mco.pdf}}}
+      \caption{Численные оценки множества Парето, полученные после 2500 испытаний}
+      \label{fig:mco_pareto}
+  \end{figure}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Тестовые многомерные задачи с ограничениями}
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+      Тестовые наборы были получены с помощью системы GCGen, позволяющей
+      сторить многомерные задачи с ограничениями из заданных функций, и нескольких
+      известных генераторов тестовых задач (GKLS, \(F_{GR}\)).
+      Характеристики каждого из сгенерированных наборов задач:
+      \begin{itemize}
+        \item \(q\)=100;
+        \item Размерность 2 или 3;
+        \item Базовые функции только GKLS или \(F_{GR}\), или их комбинация;
+        \item Количество ограничений: 2.
+      \end{itemize}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \centerline{\includegraphics[width=0.9\textwidth]{4.png}}
+    \end{column}
+  \end{columns}
+  \footnotesize{Система GCGen доступна по ссылке \textit{https://github.com/UNN-ITMM-Software/GCGen}}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Программное и техническое обеспечение}
+  \begin{center}
+    Реализация параллельного метода была выполнена на языке C++ с использованием технологии OpenMP
+    для распареллеливания процесса проведения испытаний на общей памяти.
+
+    Все вычислительные
+    эксперименты проведены на машине со следующей конфигурацией: Intel Core i7-7800X (6 cores), 64GB RAM, Unubtu 16.04 ОS, GCC 5.5 compiler.
+  \end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Результаты экспериментов на наборах синтетических задач}
+  \begin{figure}[ht]
+    \vspace*{-0.5cm}
+      \centering
+      \subfloat[\(D_{max}\)]{
+      \label{fig:max_dev} {\includegraphics[width=.5\textwidth]{mixed_2d_max.pdf}}}
+      \subfloat[\(D_{avg}\)]{
+      \label{fig:avg_dev} {\includegraphics[width=.5\textwidth]{mixed_2d_avg.pdf}}}
+      \caption{Динамика величин средней по задачам и максимальной точности в процессе решения множества двухмерных задач,
+      порождённых двумя разными генераторами GKLS и \(F_{GR}\)}
+      \label{fig:devs_mixed}
+  \end{figure}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Результаты экспериментов на наборах синтетических задач}
+  \begin{table}
+    \centering
+    \begin{tabular}{c|c|c|c|c|c}
+      %\cline{1-8}\noalign{\smallskip}
+      Класс задач & \textit{p} & Количество итераций & Время, с & \(S_i\) & \(S_t\)   \\
+      %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
+      \hline
+      GKLS \& \(F_{GR}\) based \
+        & 1 & 51434 & 90.20 & -    & - \\
+        & 2 & 25698 & 56.96 & 2.00 & 1.58 \\
+        & 4 & 13015 & 36.67 & 3.95 & 2.46 \\
+        & 6 & 8332  & 26.85 & 6.17 & 3.36 \\
+      \hline
+      GKLS based 2d \
+        & 1 & 59066 & 97.53 & -    & - \\
+        & 2 & 29060 & 60.56 & 2.04 & 1.61 \\
+        & 4 & 14266 & 38.92 & 4.14 & 2.51 \\
+        & 6 & 9436  & 29.53 & 6.26 & 3.30 \\
+      \hline
+      GKLS based 3d \
+        & 1 & 782544 & 1117.55 & -    & - \\
+        & 2 & 397565 & 752.92  & 1.97 & 1.48 \\
+        & 4 & 208073 & 526.67  & 3.76 & 2.12 \\
+        & 6 & 142089 & 445.45  & 5.50 & 2.51 \\
+      \hline
+    \end{tabular}
+  \end{table}
 \end{frame}
@@
   \frametitle{Заключение}
+  В ходе работы были достигнуты следующие результаты:
     \begin{itemize}
-      \item
+      \item реализована поддержка нелинейных ограничений в алгоритме, решающeм
+      множество задач глобальной оптимизации;
+      \item проведены численные эксперименты, демонстрирующие преимущество рассматриваемого подхода в скорости сходимости на всём множестве задач в среднем над решением задач по отдельности;
+      \item показана эффективность совместного решения множества задач на примере решения многокритериальной задачи с нелинейными ограничениями.
+    \end{itemize}
+
+    В ходе дальнейшей работы планируется:
+    \begin{itemize}
+      \item улучшить текущую реализацию алгоритма, сократив расходы на содержание поисковой информации для множества задач и тем самым улучшив
+      показатели параллельного ускорения по времени;
+      \item реализовать версию рассматриваемого алгоритма, работающего на распределенной памяти.
     \end{itemize}
 \end{frame}
+
 {
@@
   \begin{center}
-    \Large{Contacts:}
+    \Large{Контакты:}
 \vspace{0.5cm}

### 2020-05-28 2020/PAVT_Sovrasov/slides/slides.tex

@@
           \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
-          D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, \\ 1\leq{i}\leq{N},\:g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}
+          D=\{y\in \mathbb{R}^N:a_i\leqslant y_i\leqslant{b_i}, \\ 1\leqslant{i}\leqslant{N},\:g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}
         \end{array}
@@
       \begin{displaymath}
-        |f(y_1)-f(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
+        \begin{array}{cr}\\
+          |f(y_1)-f(y_2)|\leqslant L\Vert y_1-y_2\Vert,\\y_1,y_2\in \mathbb{R}^N:a_i\leqslant y_i\leqslant{b_i},
+        \end{array}
       \end{displaymath}
@@
 \begin{frame}
-  \frametitle{Редукция размерноти}
+  \frametitle{Редукция размерности}
   \begin{center}
@@
   \begin{gather}
-    \lbrace y\in \mathbb{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\} \nonumber \\
+    D_e=\lbrace y\in \mathbb{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\} \nonumber \\
     \min\{\varphi(y): y\in D\}=\min\{\varphi(y(x)): x\in [0,1]\} \nonumber
   \end{gather}
-  \(y(x)\) является негладкой кривой, отображающей отрезок \([0,1]\) на гиперкуб \(D\).
+  \(y(x)\) является негладкой кривой, отображающей отрезок \([0,1]\) на гиперкуб \(D_e\).
   \begin{figure}[ht]
@@
   \begin{displaymath}
-    \varphi(y,\lambda_1,\lambda_2)=\max\{\lambda_1 f_1(y), \lambda_2 f_2(y):\lambda_1,\lambda_2\in[0,1],\: \lambda_1+\lambda_2=1\}
+    \varphi(y^*(\lambda_1,\lambda_2))=\min_{y\in D}\max\{\lambda_1 f_1(y), \lambda_2 f_2(y)\};\lambda_1,\lambda_2\in[0,1],\: \lambda_1+\lambda_2=1
 \end{displaymath}
@@
     \begin{array}{cr}\\
-      SP(S)=\sqrt{\frac{1}{|S|-1} \sum_{i=1}^{|S|} (\overline{d}-d_i)},\:\overline{d}=mean\{d_i\}, \\
+      SP(S)=\sqrt{\frac{1}{|S|-1} \sum_{i=1}^{|S|} (\overline{d}-d_i)^2},\:\overline{d}=mean\{d_i\}, \\
       d_i=\min_{s_i,s_j\in S:s_i\ne s_j}||F(s_i)-F(s_j)||_1,\: F=(f_1,f_2)
@@
     Все вычислительные
-    эксперименты проведены на машине со следующей конфигурацией: Intel Core i7-7800X (6 cores), 64GB RAM, Unubtu 16.04 ОS, GCC 5.5 compiler.
+    эксперименты проведены на машине со следующей конфигурацией: Intel Core i7-7800X (6 cores) CPU, 64GB RAM, Unubtu 16.04 ОS, GCC 5.5 compiler.
   \end{center}
@@
       \hline
-      GKLS \& \(F_{GR}\) based \
+      GKLS \& \(F_{GR}\)-based \
         & 1 & 51434 & 90.20 & -    & - \\
@@
       \hline
-      GKLS based 2d \
+      GKLS-based 2d \
         & 1 & 59066 & 97.53 & -    & - \\
@@
       \hline
-      GKLS based 3d \
+      GKLS-based 3d \
         & 1 & 782544 & 1117.55 & -    & - \\

### 2020-05-29 2020/JPDC_MCO/2020_JPDC.tex
статья JPDC MCO 2020

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\usepackage{adjustbox}
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\journal{Journal of Parallel and Distributed Computing}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Parallel Solving of Multiple Information-Coordinated Global Optimization Problems}

%% Group authors per affiliation:
%\author{Victor Gergel \and Evgeniy Kozinov}
%\address{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia}
%\ead{gergel@unn.ru}

\author[mymainaddress]{Victor Gergel}
\ead{gergel@unn.ru}

\author[mymainaddress]{Evgeniy Kozinov}
\ead{evgeny.kozinov@itmm.unn.ru}

\address[mymainaddress]{Institute of Informational Technologies, Mathematics and Mechanics, Mathematical Center, Lobachevsky State University of Nizhny Novgorod, Nizhni Novgorod, Russia}

\begin{abstract}
This paper proposes an efficient approach for the parallel solution of computationally time consuming problems of multiple global optimization problems, in which minimized functions can be multiextremal and calculating function values may require huge amounts of computations. The proposed approach is based on the information-statistical theory of global optimization, at which a general computational scheme of global optimization methods is proposed. In the paper, this general scheme is expanded by the possible reuse of search information obtained in the process of computations when solving multiple global optimization problems. Within the framework of the proposed generalized scheme, parallel algorithms are proposed for computational systems with shared and distributed memory. Results of computational experiments demonstrated that the proposed approach can significantly reduce the computational complexity of solving multiple global optimization problems.
\end{abstract}

\begin{keyword}
Global optimization\sep dimensionality reduction\sep search information\sep parallel methods of global search\sep computational complexity.
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}\label{sec:1}

Global optimization (GO) problems are among the most complex optimization problems. In contrast to local optimization problems, when searching for a global minimum, the numerical method must evaluate the behavior of the optimized function in the entire global search area. In general, GO problems are subject to the ``dimensionality curse'', when the volume of computations performed increases exponentially with increasing dimensionality. Such computational complexity, as well as high practical demand, have led to a great deal of research activity in the field of global optimization. As a result of this research, numerous effective methods for solving global optimization problems and examples of solving practical problems have been proposed -- see, for example, \cite{c1,c2,c3,c4,c5,c6,c7,c8,c9,c10,c11,c12,c13,c14,c18}.

The paper proposes a further complication in formulating the problem, when not only one, but several global optimization problems, need to be solved simultaneously. Such problems are encountered, for example, when one needs to find several efficient solutions of multicriteria optimization problems or when part of the variable parameters in the optimization problem can take only a certain number of discrete values. The formulation of such multiple global optimization (MGO) problems is relatively new; for the first time, this type of problem was considered in \cite{c15,c16,c17}. The limited attention given to MGO problems is most likely due to the high computational complexity involved.

In general, MGO problems can be solved by a simple sequential (or parallel) solution of individual problems from a set. However, this approach means the computational complexity of solving problems increases many times. Meanwhile, MGO problems may have a certain informational interrelation -- a similar connectivity is observed, for example, when MGO problems are generated to solve the same multicriteria optimization problem -- see, for example, \cite{c19,c20}. Such interconnectedness of MGO problems can be used to reduce significantly the amount of computation performed when solving the entire set of problems.

The approach proposed in this paper assumes that the computational values of any function of the solvable set of MGO problems can be reduced to the values of any other functions of the set without any time consuming calculations. In such cases, all the search information obtained in solving a particular problem can be used in solving all other problems in the set. Such repeated reuse of search information can significantly reduce the amount of computation performed, up to just a few iterations when solving the next MGO problems. In addition, the presence of informational connectivity significantly expands the possibilities of a parallel solution for MGO problems. The paper proposes two parallel computational schemes: a cooperative scheme, which means the search information between the problems is exchanged when solving several MGO problems in parallel, and a competitive scheme, which occurs when MGO problems solved in parallel compete with each other for the use of available computational resources.

The structure of the paper is as follows. Chapter \ref{sec:2} sets out the problem of multiple global optimization. Chapter \ref{sec:3} presents a general computational scheme for solving global search problems. Chapter \ref{sec:4} proposes parallel methods for solving MGO problems by using computational systems with shared and distributed memory. Chapter \ref{sec:5} contains the results of numerical experiments confirming the effectiveness of the proposed parallel methods. In the conclusion, results are discussed and the main directions for further research are presented.

\section{Problem Statement}\label{sec:2}

The multiple global optimization (MGO) problem consists in solving a family of problems:
\begin{equation}\label{eq:1}
f_i(y^*) = \argmin f_i(y), y \in D, 1 \leq i \leq s,
\end{equation}

\begin{equation}\label{eq:2}
D  = \{ y\in R^N: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}
\end{equation}
defined by a set of optimized functions
\begin{equation}\label{eq:3}
F(y) = \{ f_1(y),  f_2(y),\dots, f_s(y) \}
\end{equation}
where $y = (y_1,y_2,\dots,y_N)$ is the vector of varied parameters, $N$ is the dimension of the global optimization problems to be solved, and $D$ is the search domain for the given vectors $a$ and $b$. 
In this paper, it is assumed that the functions $f_i(y)$, $1 \leq i \leq s$, can be multiextremal, and the procedures for calculating the values of these functions at points in the search domain $y \in D$ can be computationally expensive. It is also assumed that the functions $f_i(y)$, $1 \leq i \leq s$, satisfy the Lipschitz condition
\begin{equation}\label{eq:4}
|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, y',y''\in D, 1 \leq i \leq s
\end{equation}
where $L_i$ is the Lipschitz constant for the function $f_i(y)$, $1 \leq i \leq s$ and ${\|*\|}$ denotes the Euclidean norm in $R^N$. Fulfillment of the Lipschitz condition allows us to construct numerical estimates of the possible behavior of minimized functions based on a finite set of calculated function values at points in the search domain.

A key condition of this approach is the assumption that any calculated value of any function $f_i(y)$, $1 \leq i \leq s$, of the set $F$ at an arbitrary point $y \in D$ of the search domain $D$ can be transformed to the value of any other function of the set $F$ at the same point $y \in D$ without any time consuming computations, in a time significantly shorter compared to the computation time values of the functions $f_i(y)$, $1 \leq i \leq s$. In other words, there must be a transformation $Z$ such that

\begin{equation}\label{eq:5}
\begin{matrix}
\forall 1 \leq i, j \leq s, \bar{y} \in D \Rightarrow z_i=Z(z_j ), z_i = f_i (\bar{y}),z_j=f_j (\bar{y}), \\
T(Z) \ll T(f_l (y) ),1 \leq l \leq s,
\end{matrix}
\end{equation}
where $T(*)$ is the computation time of the expression given in parenthesis.

This condition is fulfilled, in part, when solving multicriteria optimization problems (MCO)

\begin{equation}\label{eq:6}
g(y) = (g_1(y), g_2(y), \dots , g_s(y)) \to min,  y\in D.
\end{equation}

To find efficient (unimprovable) solutions, a widely used approach is based on the reduction (scalarization) of the vector criterion $g(y)$ to the general scalar criterion, the minimization of which will lead to the Pareto optimal solution of the MCO problem \cite{c19,c20}. So, when using mini-max convolution, the vector criterion $g(y)$ from (\ref{eq:6}) reduces to the scalar criterion
\begin{equation}\label{eq:7}
\begin{matrix}
G(\lambda,y)=\max{(\lambda_i g_i(y),1 \leq i \leq s)},	\\
\lambda=(\lambda_1,\lambda_2,\dots,\lambda_s)\in \Lambda \subset R^s:\sum_{i=1}^s{\lambda_i =1},\lambda_i\geq 0,1 \leq i \leq s.
\end{matrix}
\end{equation}

If it is necessary to find several efficient decisions to the MCO problem (\ref{eq:6}), the optimization of the scalar criterion $G(\lambda,y)$ from (\ref{eq:7}) is performed for different values of the coefficients $\lambda \in \Lambda$ of the convolution of the criteria $g_i(y)$, $1 \leq i \leq s$. Thus, the problem of finding a set of efficient decisions for the MCO problem can be represented as a statement of the MGO problem (\ref{eq:1})-(\ref{eq:3}), where the optimized functions $f_i(y)$, $1 \leq i \leq s$, and the sets $F$ are defined as
\begin{equation}\label{eq:8}
f_i (y)= G(\lambda_i,y),1 \leq i \leq s.
\end{equation}

Thus, if the values of efficiency criteria $g_i(y)$, $1 \leq i \leq s$, computed at some point $y=\bar{y}$ of the search domain $D$, the value of the scalar criterion $G(\lambda,y)$ (and, accordingly, the values of the functions $f_i(y)$, $1 \leq i \leq s$, from a set $F$) can be computed at the point $\bar{y} \in D$ for any values of the coefficients $\lambda \in \Lambda$ of the convolution $g_i(y)$, $1 \leq i \leq s$ (the transformation of $Z$ from (\ref{eq:5}) is determined by the expression (\ref{eq:7}) at fixed values of the criteria $g_i(y)$, $1 \leq i \leq s$) \cite{c15,c16,c17}.

The ability to quickly convert the calculated values of optimized functions determines the need to accumulate and use of search information obtained in the calculation process

\begin{equation}\label{eq:9}
\omega_k (j)=\{(y^i,f^i=f_j(y^i)):1 \leq i \leq k \},1 \leq j \leq s,
\end{equation}
where $j$, $1 \leq j \leq s$, is the number of the current minimized function $f_j(y)$ of the set $F$ from (\ref{eq:3}), $k$ is the number of iterations of the global search, $y^i \in D$, $1 \leq i \leq k$,  are points in which the values of the optimized problems of the set $F$ were computed, and $z^i$, $1 \leq i \leq k$, are the values of the function $f_j(y)$ at these points. Global optimization methods can use this search information, thereby ensuring the ability to solve each successive problem of the $F$ set, taking into account the results of all previous calculations. Numerical experiments that were performed (see Section \ref{sec:5}), confirm that such repeated use of search information reduces the number of iterations of the global search required to solve the next problem of the set $F$ (even to the point of performing just a few search iterations).

In conclusion, it should be noted that, in general, the set of minimized functions $F$ from (\ref{eq:3}) can be formed dynamically in the course of calculations by adding new or removing existing global optimization functions

\begin{equation}\label{eq:10}
F'=F+(-) f(y).
\end{equation}

\section{A General Computational Scheme for Solving Global Optimization Problems}\label{sec:3}

The approach proposed in this paper is based on two key ideas used in the framework of the information-statistical theory of global optimization \cite{c5}.

In accordance with this theory, first of all, dimension reduction is achieved using the Peano space-filling curves (evolvents) $y(x)$, uniquely and continuously mapping the interval $[0,1]$ to the $N$-dimensional domain $D$ -- see, for example, \cite{c5,c21}. As a result of this reduction, multidimensional global optimization problems (\ref{eq:7}) can be reduced to one-dimensional ones:

\begin{equation}\label{eq:11}
\min {f_i (y(x))} = \min{f_i (y)}, x \in [0,1], y \in D, 1 \leq i \leq s.
\end{equation}

It should be noted that the one-dimensional functions obtained as a result of the reduction satisfy the uniform H\"older condition, i.e.
\begin{equation}\label{eq:12}
|f_i (y(x'))-f_i (y(x''))| \leq H_i \|x'-x''\|, x',x''\in [0,1], 1 \leq i \leq s,
\end{equation}
where the constants $H_i$ are determined by the relation $H_i = 2L_i \sqrt{N+1}$, ${1 \leq i \leq s}$, $L_i$ are the Lipschitz constants from (\ref{eq:4}), and $N$ is the dimensionality of the optimization problem (\ref{eq:1}).

Dimensionality reduction significantly decreases the computational complexity of analyzing multidimensional data to highlight the subdomains of the search domain $D$ for the efficient continuation of the global search, and reduces the severity of the ``curse of dimensionality'' problem in solving multidimensional optimization problems. After dimensionality reduction, the search information $\omega_k$ from (\ref{eq:9}) takes the form

\begin{equation}\label{eq:13}
A_k (j)=\{ (x_i, y(x_i), z_i=f_j (y(x_i) ): 1 \leq i \leq k,1 \leq j \leq s \},
\end{equation}
in which, in addition to the reduced representation, the data is arranged in the ascending order\footnote{Data ordering is reflected by using a subscript.} of points $x_i$,  $1 \leq i \leq k$, i.e.

\begin{equation}\label{eq:14}
x_1< x_2< \cdots < x_k
\end{equation}
for a more efficient execution of global search algorithms. It should be noted that in optimizing the reduced one-dimensional functions $f_i (y(x))$, $1 \leq i \leq s$, many one-dimensional global search algorithms can be used (possibly after some generalization).

Numerical methods for constructing approximations of Peano space-filling curves (evolvents) with a given accuracy for a given dimensionality are considered in \cite{c5}; examples of such approximations are shown in Fig. \ref{fig:1}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig1}
  \caption{Examples of approximations of the Peano space-filling curves (the approximation density is equal to 2, 3 and 4 correspondingly) \cite{c5}}
  \label{fig:1}
\end{figure}

Further, in the framework of information-statistical theory, a general computational scheme of global optimization algorithms is proposed. This scheme can be summarized as follows.

Let $k$, $k \geq 2$ optimization iterations be performed, and the search information $A_k$ from (\ref{eq:13}) is obtained. To adaptively select the points of the next iterations, the optimization algorithm has to evaluate the possibility of placing the global minimum in the subintervals, into which the interval $[0,1]$ is divided by the points $x_i$, $1 \leq i \leq k$, previously performed iterations of the global search. This assessment can be performed by introducing characteristics of the subintervals, the value of which would be proportional to the degree to which the global minimum can be located in these subintervals

\begin{equation}\label{eq:15}
R(i)=\mathcal{R} (i,A_k ), 1 < i \leq k.
\end{equation}

The type of characteristics $R(i)$, $1 < i \leq k$, is a key design element of global optimization algorithms. For example, for an algorithm that builds a uniform dense grid in the search domain, the characteristics may simply be the length of the interval

\begin{equation}\label{eq:16}
R(i)=(x_i-x_{i-1}), 1 < i \leq k, x_i, x_{i-1} from A_k.
\end{equation}

For the multidimensional generalizations of the algorithms proposed in \cite{c22,c23}, the characteristic is an estimate of the minimum possible value of the minimized function on the interval, i.e.

\begin{equation}\label{eq:17}
R(i)=0.5 H \rho_i - 0.5 (z_{i-1} + z_i), \rho_i=(x_i - x_{i-1})^{1/N}, 1 < i \leq k,
\end{equation}
where $H$ is the H\"older constant from (\ref{eq:12}) for the reduced global optimization problem to be solved; ($N$ is the dimensionality of the problem from (\ref{eq:7})). For the multidimensional generalized global search algorithm (GSA) \cite{c5,c24}, developed in the framework of the information-statistical approach, the characteristic has the form

\begin{equation}\label{eq:18}
R(i)=m \rho_i+\frac{(z_i-z_{i-1})^2}{m \rho_i }-2(z_{i-1}+z_i ), \rho_i=(x_i-x_{i-1})^{1/N}  ,1 < i \leq k,
\end{equation}
where $m$ is the numerical estimate of the H\"older constant obtained from the available search information $A_k$ from (\ref{eq:13})

\begin{equation}\label{eq:19}
m=r M, M=\frac{\max|z_i-z_{i-1}|}{\rho_i}, \rho_i=(x_i-x_{i-1})^{1/N}  ,1 < i \leq k
\end{equation}
($r>1$ is the parameter of the GSA algorithm).

The presence of interval characteristics makes it possible to represent the procedure for iterating a global search as the following sequence of steps \cite{c5}.

\textit{Step 1.} Calculate the characteristics of the intervals $R(i)$, $1 < i \leq k$, and determine the interval with the maximum characteristic
\begin{equation}\label{eq:20}
R(t)=\max{R(i)}, 1 < i \leq k.
\end{equation}

\textit{Step 2.} Select the next iteration point in the interval with the maximum characteristic \footnote{The rule $X$ for choosing the point of the next iteration in the interval $(x_{t-1},x_t)$ is another design element of characteristically represented algorithms.}
\begin{equation}\label{eq:21}
x^{k+1}=X(x_{t-1},x_t ).
\end{equation}

Calculate the value $z^{k+1}$ of the minimized function at this point (the procedure for calculating the value of the function will be called a trial) and add the obtained data $(x^{k+1},z^{k+1})$ to the search information $A_k$ from (\ref{eq:13}).

\textit{Step 3.} Check the stopping condition 
\begin{equation}\label{eq:22}
\rho_t \leq \varepsilon,	
\end{equation}
where $\rho_t=(x_t-x_{t-1})^{1/N}$, $t$ from (\ref{eq:20}) and $\varepsilon > 0$ is the specified accuracy of the solution to the optimization problem. If the stopping condition (\ref{eq:22}) is reached, then the solution to the optimization problem is stopped, otherwise $k=k+1$ is assumed and the next iteration of the global search is performed.

After the calculations are completed, the lowest calculated value of the minimized function can be taken as the global minimum estimate
\begin{equation}\label{eq:23}
z_k^*=\min{z_i}, 1 \leq i \leq k.
\end{equation}

Again, we note that the above computational scheme is quite general. Within the framework of this scheme, many global search algorithms can be presented, as evidenced, in particular, by the examples given in (\ref{eq:16})-(\ref{eq:18}), as well as other characteristically represented algorithms (see, for example, \cite{c22,c23,c24,c25,c26,c27,c28,c29,c30,c31}). Thus, the parallelization methods discussed in Section \ref{sec:4} can be used for a sufficiently wide variety of global search algorithms.

The convergence conditions for the characteristically represented algorithms depend on the properties of the interval characteristics used. For example, one of the sufficient conditions for convergence of the algorithms is the requirement that, when iterating the global search, the characteristic of the interval containing the global minimum point repeatedly takes the maximum value in Step 1 of the general characteristic scheme. This condition is fulfilled, for example, for the multidimensional generalizations of the algorithms proposed in \cite{c22,c23} for the exact setting of the H\"older constant in (\ref{eq:17}). For GSA, the sufficient condition for convergence is the relation \cite{c5}
\begin{equation}\label{eq:24}
m \geq 2H,
\end{equation}
starting with some iteration of the global search. Additional information on the convergence of characteristically represented global search algorithms can be obtained, for example, in \cite{c5}.

In the framework of the proposed approach, the GSA algorithm described by rules (\ref{eq:18})-(\ref{eq:23}) will be used to solve global optimization problems of the set $F$. Moreover, due to the informational connectivity (\ref{eq:5}) of the problems of the set $F$, the GSA algorithm can use all the search information $A_k$ from (\ref{eq:13}) when solving the next global optimization problem. The GSA algorithm for solving problems of the set $F$ and using search information $A_k$, will be referred to as the multiple global search algorithm (MGSA).

\section{Parallel Computations for Solving Multiple Global Optimization Problems}\label{sec:4}

In keeping with the initial assumption that it is computationally expensive to calculate the values of minimized functions of the set $F$ from (\ref{eq:3}), solving all global optimization problems (\ref{eq:1}) may require a huge amount of calculations, which increases with the number of problems to be solved. Thus, obtaining estimates of globally optimal decisions with reasonable time delays can only be achieved using parallel high-performance computational systems.

An obvious way to apply parallel computing is to solve the problems of the set $F$ in parallel. This method is simple and does not require any additional efforts for implementation. With this approach and with a sufficient number of computational devices (cores / processors), the time needed to solve all the $F$ set will be determined by the time used in solving the most time-consuming problem. However, this time may turn out to be quite long and necessitate the use of parallel computing to solve even one global optimization problem.

Unfortunately, well-known and widely used parallelization methods are poorly applicable in the case of global optimization problems. For example, search domain distribution between processors for global optimization problems leads to the global minimum of the problems to be solved being located on only one of the processors; in this case, all other processors will perform redundant calculations.

In the framework of the information-statistical theory, a general approach to parallelizing global optimization problems has been proposed -- parallel computing is provided by parallel (simultaneous) computing of the values of the minimized function at several different points in the search domain \cite{c5}. This approach is general -- it can be applied to solving any global optimization problems in a parallel mode. And such an approach is efficient, since the most time consuming part of the problem is parallelized due to the initial assumption that calculating the values of minimized functions is computationally expensive.

Next, we propose three different ways of applying this approach for parallel solutions to multiply global optimization problems (\ref{eq:1})-(\ref{eq:3}).

\subsection{Strategy 1: The Problem-oriented Parallel Scheme} \label{subsec:1}

In this method of parallel computing, it is proposed to solve the set $F$ problems strictly sequentially, and to use all available computational resources only to solve the one optimization problem. When solving problems of the set $F$ sequentially, all available search information $A_k$, obtained in the calculation process is reused.

To implement this approach, it is necessary to ensure that at each iteration of the global optimization problem, not one, but several points in the search domain $D$ are selected for parallel calculation of the values of the minimized function at these points. In keeping with the fact that characteristically represented global search algorithms use the characteristics of the intervals $R(i)$, $1 < i \leq k$, from (\ref{eq:15}), to assess the possibility of locating the global minimum in the subinterval of the interval $[0,1]$, it is most reasonable to select the points of the next iteration of the global search in the intervals with the maximum values of the characteristics \cite{c33,c34,c35}.

Performing iterations of a global search for a parallel generalization of the MGSA algorithm (PMGSA-1) can be represented as follows ($q>1$ is the number of available computational devices (cores/processors) with shared memory) \cite{c33,c34,c32}.

\textit{Step 1}. Calculate the characteristics of the intervals $R(i)$, $1 < i \leq k$, and determine $q$ intervals with the maximum values of characteristics

\begin{equation}\label{eq:25}
R(t_j)=\max{R(i)},1 < i \leq k, 1 \leq j\leq q,
\end{equation}
(for $j>1$,$t_j \neq t_l$, $1 \leq l < j$ should be satisfied).

\textit{Step 2}. Select $q$ points of the next parallel iteration in the intervals with maximum characteristics

\begin{equation}\label{eq:26}
x^{k+j}=X(x_{t_j-1},x_{t_j} ), 1 \leq j \leq q,
\end{equation}
then calculate the values of $z^{k+j}$, $1 \leq j \leq q$, of the minimized function at these points and add the resulting data $(x^{k+j},z^{k+j})$, $1 \leq j \leq q$, to the search information $A_k$ from (\ref{eq:13}).

\textit{Step 3}. Check the stopping condition

\begin{equation}\label{eq:27}
\rho_{t_j} \leq \varepsilon, 1 \leq j \leq q,
\end{equation}
where $\rho_{t_j}=(x_{t_j}-x_{t_j-1} )^{1/N}$, $t_j$, $1 \leq j \leq q$, from (\ref{eq:25}) and $\varepsilon > 0$ is the specified accuracy for solving the optimization problem. If the stopping condition (\ref{eq:27}) is reached for at least one interval $t_j$, $1 \leq j \leq q$, then the solution to the optimization problem stops; otherwise $k=k+q$ is assumed and the next parallel iteration of the global search is performed.

Results of the numerical experiments show that this approach of parallelizing the MGSA algorithm allows one to obtain near-linear speedup using up to $2^N$ computational cores ($N$ is the dimensionality of the global optimization problems to be solved). An example of using PMGSA-1 is given in Fig. \ref{fig:2}. On the left in Fig. \ref{fig:2} the MGSA result is given (55 iterations), on the right the PMGSA-1 result is shown (30 iterations, i.e., speedup is 1.8). PMGSA-1 uses 2 cores, iteration points selected by the cores are marked by the different lines.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig2}
  \caption{An example of using PMGSA-1 for solving the univariate optimization problem: a) the graph of the minimized function, b) step-by-step selection of iteration points computed by the method (the ordinate axis is the iteration number, the abscissa axis is the search domain)}
  \label{fig:2}
\end{figure}

\subsection{Strategy 2: The Competitive Parallel Scheme} \label{subsec:2}

Selecting several points at once for the same search iteration in the PMGSA-1 algorithm can result in redundant points compared to the sequential GSA version, since these points are selected at the same state of search information $A_k$ from (\ref{eq:13}). Such an undesirable effect can be reduced by reducing the number of computational devices allocated to solve an optimization problem, while simultaneously increasing the number of parallel global optimization problems that are solved due to freed computational resources. In this case, the available computational resources can be distributed dynamically among problems in the global search process.

Assume, as before, that $q>1$ is the number of available computational devices (cores/processors) with shared memory, and $ns>1$ is the number of parallel problems solved from the set $F$ from (\ref{eq:3})\footnote{Without lost of generality, we will use sequential numbering of problems solved in parallel from $1$ to $ns$.}. Each optimization problem solved in parallel will compete for available computational resources -- the distribution of resources between problems can be carried out, as before, based on the values of interval characteristics. To calculate the characteristics, the problems solved in parallel have sets of search information $A_k(j)$, $1\leq j \leq ns$, from (\ref{eq:13}), which match the set of iteration points $x_i$, $1 \leq i \leq k$, of the global search, but differ, accordingly, in the values of the minimized functions. The notation $\tau_{ij}$ will then designate the interval $t_j$ in the search information $A_k(i)$ of problem $i$.

The new parallel version of the MGSA algorithm (the PMGSA-2 algorithm) obtained as a result of this approach is determined by the following sequence of steps for performing parallel iterations of the global search.

\textit{Step 0}. For each problem to be solved in parallel, normalize the values of the minimized function in the available search information in order to compare the interval characteristics in different global optimization problems\footnote{After transformation (\ref{eq:28}), $z_k^*=0$ and $m_k=1$ are carried out for all $ns$ parallel problems.}

\begin{equation}\label{eq:28}
z(j)_i'=\frac{z(j)_i-z(j)_k^*}{m(j)_k},1 \leq i \leq k,1 \leq j \leq ns,
\end{equation}
where $z(j)_k^*$, $1 \leq j \leq ns$, are estimates (\ref{eq:23}) of the minimum values of the minimized functions, and $m(j)_k$, $1 \leq j \leq ns$, are the estimates (\ref{eq:19}) of the H\"older constant from (\ref{eq:12}). Further, when performing the steps of the algorithm, the converted values $z(j)_i'$, $1 \leq i \leq k$, $1 \leq j \leq ns$ should be used when calculating the characteristics of intervals and points of new iterations of the global search.

\textit{Step 1}. Calculate the characteristics of intervals $R(i)$, $1 < i \leq k$, for all $ns$ problems to be solved in parallel and determine $q$ intervals with the maximum values of the characteristics
\begin{equation}\label{eq:29}
R(\tau_{l(j)j})=\max{R_\nu(i)}, 1 < i \leq k, 1 \leq \nu \leq ns, 1 \leq j \leq q,
\end{equation}
where $R_\nu(i)$ is the characteristic of the interval $i$ in the search information $A_k(\nu)$ of the problem with the number $\nu$, $l(j)$ is the number of the problem to be solved, in which the maximum characteristic $R(\tau_{l(j)j})$ was chosen; (in (\ref{eq:29}) for $j>1$, $\tau_{l(j)j}\neq \tau_{l(l)l}$, $1 \leq l < j$ should be satisfied).

\textit{Step 2}. The execution of Step 2 is divided into three sequential stages: 
\begin{enumerate}

	\item Select $q$ points of the next parallel iteration in intervals with maximum characteristics
\begin{equation}\label{eq:30}
 x(l(j))^{k+j}=X(x_{\tau_{l(j)j}-1},x_{\tau_{l(j)j}}),1 \leq j \leq q	
\end{equation}
(the point $x(l(j))^{k+j}$ is determined in the interval $(x_{\tau_{l(j)j}-1},x_{\tau_{l(j)j}})$, belonging to the search information $A_k (l(j))$ of problem $l(j)$.

	\item Calculate the values $z(l(j))^{k+j}$, $1 \leq j \leq q$, of minimized functions at the points $x(l(j))^{k+j}$, $1 \leq j \leq q$.

  \item Add the obtained data $(x(l(j))^{k+j}, z(l(j))^{k+j})$, $1 \leq j \leq q$, in the search information $A_k(i)$, $1 \leq i \leq ns$, of each problem being solved in parallel with the necessary values transformation $z(l(j))^{k+j}$, $1 \leq j \leq q$, using the $Z$ transformation from (\ref{eq:5}).

\end{enumerate}

\textit{Step 3}. Verify the stopping condition
\begin{equation}\label{eq:31}
\rho_{\tau_{l(j)j}} \leq \varepsilon, 1 \leq j \leq q,
\end{equation}
where $\rho_{\tau_{l(j)j}} = (x_{\tau_{l(j)j}} - x_{\tau_{l(j)j)-1}} )^{1/N}$, $\tau_{l(j)j}$, $1 \leq j \leq q$, from (\ref{eq:29}) and $\varepsilon > 0$ is the given accuracy for solving the optimization problem. If the stopping condition (\ref{eq:31}) is reached for at least one interval $\tau_{l(j)j}$, $1 \leq j \leq q$, then the solution to the optimization problem stops; otherwise, $k=k+q$ is assumed and the next parallel iteration of the global search is performed.

Let us briefly explain the rules for performing a parallel iteration of the global search. The characteristics of the intervals are calculated for all $ns$ problems to be solved in parallel (Step 1), and $q$ intervals with the maximum values of the characteristics are selected among them. As a result, the available computational resources are distributed dynamically among problems -- computational devices can be distributed evenly between problems, or they can be allocated to only one problem with the best interval characteristics. After calculating the values of the minimized functions (Step 2), the new search data obtained is added to the search information of all solved problems; regardless of where the intervals for iterating the global search were selected, the search information of all solved problems is updated with a full set of new search data.

The maximum efficiency of the proposed PMGSA-2 algorithm is achieved if the global minimum in parallel problems is reached at the same point in the search domain. In any case, getting additional search information from simultaneously solved problems can significantly reduce the time needed to solve each global search problem -- see the results of numerical experiments in Section \ref{sec:5}. An example how PMGSA-2 selects the intervals with the maximum characteristics (Step 1) is given in Fig.\ref{fig:3}.

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{fig3}
  \caption{An example of intervals selection by the PMGSA-2 algorithm (Step 1)}
  \label{fig:3}
\end{figure}

\subsection{Strategy 3: The Cooperative Parallel Scheme} \label{subsec:3}

When using computational devices with shared memory, the number of cores / processors is quite limited -- the number of processors on one computational node is usually 2, the number of computational cores in one processor usually does not exceed 10-20. For increasing possible parallelism, it becomes inevitable to use computational nodes where memory is distributed.

Let $p>1$ be the number of computational nodes with distributed memory, each of which has $q>1$ computational cores (that is, the total number of cores in a computational system is $p*q$).

In the framework of the proposed approach, a parallel multiple search algorithm (the PMGAS-3 algorithm) is proposed for distributed memory computational systems, the main rules of which are as follows.

1) Solving the problems of the set $F$ from (\ref{eq:3}) is carried out sequentially in groups; the number of problems in each group coincides with the number of available computational nodes.

2) Each problem of the current group is solved on its own computational node using the PMGSA-1 algorithm.

At the end of each parallel iteration of the global search, the results of the iteration $(x^{k+j}, z^{k+j})$, $1 \leq j \leq q$, of each processor are transferred to all available nodes of the computational system with the necessary transformation of the values of minimized functions in accordance with the transformation $Z$ from (\ref{eq:5}).

Before performing a new iteration, each computational node accepts the search data transmitted to it and adds this data to the search information $A_k$ from (\ref{eq:13}).

In this way, in keeping with the above rules, the PMGSA-3 algorithm is distributed -- the problems of the set $F$ are solved independently on different computational nodes. During the global search process, computational nodes exchange the calculation results that are obtained; after each iteration of the global search, the search information for each problem being solved is updated with $q$ calculated values of the minimized function. Thus, the more computational nodes there are (and accordingly, simultaneously solved problems), the faster as a rule, each optimization problem is solved. This is evidenced, in particular, by the results of numerical experiments that were performed (see Section \ref{eq:5}). An example how PMGSA-3 performs an optimization iteration is given in Fig. \ref{fig:4}.

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{fig4}
  \caption{Intervals and trial points selected by the PMGSA-3 algorithm}
  \label{fig:4}
\end{figure}

\section{Results of Numerical Experiments}\label{sec:5}

Numerical experiments were performed on the supercomputers Lobachevsky (University of Nizhni Novgorod), Lomonosov (Moscow State University), MVS-10P (Joint Supercomputer Center of RAS) and supercomputers Endeavor. The numerical results were obtained by using the following computational nodes: 2 Intel Xeon Platinum 8260L, 2.4 GHz, 256 GB RAM (i.e. a total of 48 CPU cores were available on each node). The executable program code was built by using the Intel Parallel Studio XE 2019 software package. The numerical experiments were performed using the Globalizer system \cite{c40}.

First, we present the results achieved when comparing the MGSA algorithm with several multicriteria optimization algorithms \cite{c37}. In carrying out the experiments, the two-criteria test problem proposed in \cite{c36} was used:

\begin{equation}\label{eq:32}
f_1(y)=(y_1-1) y_2^2+1,f_2 (y)=y_2, 0 \leq y_1, y_2 \leq 1.
\end{equation}

Solving the MCO problem is understood to mean the construction of a numerical approximation of the Pareto domain. To assess the quality of the approximation, the completeness and uniformity of coverage of the Pareto domain were compared using the following two indicators \cite{c15,c36}:
\begin{itemize}
	 \item The hypervolume index (HV). This indicator characterizes how completely the Pareto domain is approximated (a higher value corresponds to a more complete coverage of the Pareto domain). 
	 \item The distribution uniformity index (DU). This indicator characterizes how uniformly the Pareto domain is covered (a lower value corresponds to a more uniform coverage of the Pareto domain).
\end{itemize}

During the experiments, five multicriteria optimization algorithms were compared: the Monte-Carlo (MC) method, the genetic algorithm SEMO from the PISA library \cite{c37,c38}, the Non-uniform coverage (NUC) method \cite{c36}, the bi-objective Lipschitz optimization (BLO) method \cite{c37} and the MGSA algorithm. For the MGSA  algorithm, 50 subproblems $G(\lambda, y)$ from (\ref{eq:32}) were solved for different values of convolution coefficients $\lambda$, uniformly distributed in the interval $[0,1]$. The results of the experiments are presented in Table \ref{tab:1}.

\begin{table}[ht]
\centering
\caption{Comparison of the effectiveness of multicriteria optimization algorithms \cite{c37})}
\label{tab:1}
\begin{tabular}{cccccc}
\hline
Method                                                                                     & MC    & SEMO  & NUC   & BLO   & \textbf{MGSA} \\ \hline
Iteration number                                          & 500   & 500   & 515   & 498   & \textbf{370}   \\
\begin{tabular}[c]{@{}c@{}}Number of Pareto optimal points\end{tabular} & 67    & 104   & 29    & 68    & \textbf{100}    \\
HV                                                        & 0.300 & 0.312 & 0.306 & 0.308 & \textbf{0.316} \\
DU                                                        & 1.277 & 1.116 & 0.210 & 0.175 & \textbf{0.101} \\ \hline
\end{tabular}
\end{table}

These numerical results show that the MGSA algorithm has a noticeable advantage in almost all performance indicators when compared with the methods of multicriteria optimization being considered, even when solving relatively simple MCO problems.

Further the numerical experiments were performed for evaluating the efficiency of the parallel PMGSA-1,2,3 algorithms in solving the MGO problems. In the experiments, the MGO problems were generated as problems seeking a set of efficient solutions to the MCO problems using the minimax convolution $G (\lambda, y)$ from (\ref{eq:7}) for different values of the convolution coefficients $\lambda$, uniformly distributed in the domain $\Lambda$. Given the initial assumption about the high complexity of computing the values of minimized functions of the set $F$ from (\ref{eq:3}), the efficiency of the proposed methods is evaluated by the number of optimization iterations performed by the methods to achieve the required accuracy under the stopping condition from (\ref{eq:22}).

In the first series of numerical experiments, five-criteria four-dimensional MCO problems were solved, i.e. $N = 4$, $m = 5$. In the experiments, 100 MCO problems were solved, in which multiextremal functions obtained by using the GKLS generator \cite{c39} were used as the efficiency criteria; for each problem 100 efficient solutions were calculated (i.e., a total of $10,000$ subproblems $G (\lambda, y)$ were solved). The results obtained (the number of search iterations performed before the stopping condition was fulfilled) were averaged out over the number of MCO problems solved. In solving the problems, the search accuracy $\varepsilon = 0.05$ from (\ref{eq:22}) and the reliability of the method $r = 5.6$ from (\ref{eq:19}) were used.

The results of the numerical experiments are presented in Table \ref{tab:2}. The first three rows of the table contain information about the computational resources used during the experiments. The first row marked as ``$p$'' indicates the number of computational nodes used. The number of cores used within a single node is presented in the second row marked as ``$q$''. The total number of cores used is presented in the third row marked as ``$p*q$''. The rows four through six show the average number of iterations performed by PMGSA-1,2,3 for solving a single MCO problem accordingly.

\begin{table}[ht]
\centering
\caption{The results of a series of experiments to solve five-criteria four-dimensional MCO problems}
\label{tab:2}
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{cccccccc}
\hline
\multicolumn{8}{c}{Computational resources used}                                                                                                                           \\ \hline
p                   & 1                     & 1                   & 1                  & 1                  & 2                  & 5                  & 10                 \\
q                   & 1                     & 5                   & 25                 & 48                 & 48                 & 48                 & 48                 \\
p*q                 & 1                     & 5                   & 25                 & 48                 & 96                 & 240                & 480                \\ \hline
\multicolumn{8}{c}{Average number of iterations for solving one   MCO problem}                                                                                             \\ \hline
PMGSA-1             & 117,490.8             & 19,586.9            & 4,232.4            & 2,116.5            & -                  & -                  & -                  \\
PMGSA-2             & 130,027.1             & 20,002.0            & 4,320.1            & 2,160.7            & -                  & -                  & -                  \\
PMGSA-3             & 117,490.8             & 19,586.9            & 4,232.4            & 2,116.5            & 1,064.6            & 534.3              & 285.5              \\ \hline
\multicolumn{8}{c}{Speedup obtained by using parallel methods}                                                                                                             \\ \hline
PMGSA-1             & 1                     & 6.0                 & 27.8               & 55.5               & -                  & -                  & -                  \\
PMGSA-2             & 1                     & 6.5                 & 30.1               & 60.2               & -                  & -                  & -                  \\
PMGSA-3             & 1                     & 6.0                 & 27.8               & 55.5               & 110.4              & 219.9              & 411.5              \\ \hline
\multicolumn{8}{c}{\begin{tabular}[c]{@{}c@{}}Overall reduction of executed iterations provided   \\ by parallel computations and reusing search information\end{tabular}} \\ \hline
PMGSA-1             & 11.9                  & 71.3                & 330.0              & 659.9              & -                  & -                  & -                  \\
PMGSA-2             & 10.7                  & 69.8                & 323.3              & 646.5              & -                  & -                  & -                  \\
PMGSA-3             & 11.9                  & 71.3                & 330.0              & 659.9              & 1,312.0            & 2,614.4            & 4,892.5            \\ \hline
\end{tabular}
\end{adjustbox}
\end{table}

It should be noted that for the initial sequential GSA algorithm without reusing search information, an average of 1,396,804.1 iterations are required to solve a single MCO problem. The overall reduction of the number of executed iterations, taking into account the effect of reusing search information, is indicated in the last three rows of Table \ref{tab:2}.

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{fig5}
  \caption{Speedup and efficiency of the PMGSA-3 algorithm in solving five-criteria four-dimensional MKO problems}
  \label{fig:5}
\end{figure}

It follows from the results of the experiments (see also Fig. \ref{fig:5}), that the proposed parallel methods provide superlinear speedup\footnote{Superlinear speedup can be achieved due to the fact that the sets of iteration points that are selected when performing iterations of the global search by sequential and parallel algorithms are not identical (the search information of the methods may differ).} and high scalability (maintaining the efficiency of parallel computing with an increase in the number of computational cores used). When using a single computational node and all available computational cores for shared memory computational systems, the computational speedup exceeds 55 (with 48 computational cores in use). When using 10 computational nodes (480 cores) with distributed memory, the speedup becomes 411.5. Taking into account the effect of search information reuse, the overall reduction of the number of global optimization iterations is more than 4,890 times.

In the second series of the experiments, the problems being solved were made more complicated by increasing the dimensionality -- the five-criterion six-dimensional MCO problem was solved, i.e., $N = 6$, $m = 5$. For solving these problems the parallel PMGSA-3 method was applied. When performing the experiments, the accuracy of the method $\varepsilon = 0.1$ from (\ref{eq:22}) and the reliability of the method $r = 5.6$ were used. The results of the experiments are presented in Table \ref{tab:3}.

The first three rows of the table, as before, present the number of computational nodes used, the number of cores per a node and the total number of computational cores. The last three rows correspond to the number of global search iterations performed, the speedup achieved, and the overall reduction of executed iterations, taking into account the reuse of information (the average number of iterations performed by the sequential GSA algorithm without the reuse of search information for solving one MCO problem is, on average, 3,131,284.8).
Based on the results of the experiments, it can be seen that when the dimensionality of the problem is increased, the efficiency of parallel computations using the PMGSA-3 algorithm also increases (when using 480 cores, the speedup is 472.9). Taking into account the effect of search information reuse, the number of global optimization iterations is reduced, in total, by more than 4,160 times.

\begin{table}[ht]
\centering
\caption{Results of experiments to solve the five-criterion six-dimensional MCO problem using the parallel PMGSA-3 method}
\label{tab:3}
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{ccccccc}
\hline
\multicolumn{7}{c}{Computational resources used}                                                                                                  \\ \hline
p                                                                                & 1         & 1        & 1        & 1        & 5       & 10      \\
q                                                                                & 1         & 5        & 25       & 48       & 48      & 48      \\
p*q                                                                              & 1         & 5        & 25       & 48       & 240     & 480     \\ \hline
\multicolumn{7}{c}{Results of numerical experiments}                                                                                              \\ \hline
Iterations                                                                       & 364,713.2 & 97,387.5 & 19,791.0 & 11,294.0 & 1,561.2 & 771.2   \\
Speedup                                                                          & 1         & 3.7      & 18.4     & 32.3     & 233.6   & 472.9   \\
\begin{tabular}[c]{@{}c@{}}Overall reduction \\ of iteration number\end{tabular} & 8.6       & 32.2     & 159.6    & 277.5    & 2,025.6 & 4,160.1 \\ \hline
\end{tabular}
\end{adjustbox}
\end{table}

\section{Conclusion}\label{sec:6}

The paper considers a new class of optimization problems -- multiple global optimization problems, in which minimized functions can be multiextremal, and calculating function values may require a huge amount of calculations. In addition, the set of problems to be solved can change dynamically during calculations by adding or removing solvable global optimization problems. Problems of this kind are encountered, for example, when it is necessary to find several efficient solutions of multicriteria optimization problems or when part of the variable parameters of optimization problems can take only a number of discrete values. Problems of this class have a high computational complexity and the ability to solve such problems efficiently can be achieved only by using high-performance computational systems.

The proposed approach assumes that information connectivity exists in the global optimization problems being solved, when the calculated values of any function of the set of problems to be solved can be transformed to the values of any other functions without time consuming calculations. In such cases, all the search information obtained in solving a particular optimization problem can be used to solve all the other problems. Such reuse of search information can significantly reduce the amount of calculations performed, up to just a few iterations when solving the next problems of multiple global optimization.

The availability of informational connectivity significantly expands the possibilities of parallel solutions of the entire set of multiple global optimization problems. The paper proposes two parallel computational schemes: a cooperative scheme, when several problems are solved simultaneously, with the search information that is obtained being exchanged between problems, and a competitive scheme, when problems that are simultaneously being solved compete with each other for using the available computational resources.

Results of computational experiments show that the proposed approach allows us to significantly -- by tens and hundreds of times -- reduce the computational complexity of solving multiple global optimization problems.
In general, one can observe that the proposed approach is promising and requires further research. First, continued computational experiments must be carried out to solve the problems of multiple global optimization of varying computational complexity. It is also necessary to assess the possibility of using other global optimization algorithms within the framework of the proposed approach.

\section*{Acknowledgements} 
This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''

\bibliography{mybibfile}

\end{document}

### 2020-05-29 2020/JPDC_MCO/2020_JPDC.tex
Compress references

@@
 \begin{equation}\label{eq:3}
-F(y) = \{ f_1(y),  f_2(y),\dots, f_s(y) \}
+F(y) = \{ f_1(y),  f_2(y),\dots, f_s(y) \},
 \end{equation}
@@
 \begin{equation}\label{eq:4}
-|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, y',y''\in D, 1 \leq i \leq s
+|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, y',y''\in D, 1 \leq i \leq s,
 \end{equation}
-where $L_i$ is the Lipschitz constant for the function $f_i(y)$, $1 \leq i \leq s$ and ${\|*\|}$ denotes the Euclidean norm in $R^N$.
+where $L_i$ is the Lipschitz constant for the function $f_i(y)$, $1 \leq i \leq s$, and ${\|*\|}$ denotes the Euclidean norm in $R^N$.
 Fulfillment of the Lipschitz condition allows us to construct numerical estimates of the possible behavior of minimized functions based on a finite set of calculated function values at points in the search domain.
@@
 
-Let $k$, $k \geq 2$ optimization iterations be performed, and the search information $A_k$ from (\ref{eq:13}) is obtained.
+Let $k$, $k \geq 2$, optimization iterations be performed, and the search information $A_k$ from (\ref{eq:13}) is obtained.
 To adaptively select the points of the next iterations, the optimization algorithm has to evaluate the possibility of placing the global minimum in the subintervals, into which the interval $[0,1]$ is divided by the points $x_i$, $1 \leq i \leq k$, previously performed iterations of the global search.
@@
 \begin{equation}\label{eq:16}
-R(i)=(x_i-x_{i-1}), 1 < i \leq k, x_i, x_{i-1} from A_k.
+R(i)=(x_i-x_{i-1}), 1 < i \leq k, \; x_i, x_{i-1} \; \mathrm{from} \; A_k.
 \end{equation}
@@
 \end{equation}
-where $H$ is the H\"older constant from (\ref{eq:12}) for the reduced global optimization problem to be solved; ($N$ is the dimensionality of the problem from (\ref{eq:7})).
+where $H$ is the H\"older constant from (\ref{eq:12}) for the reduced global optimization problem to be solved; $N$ is the dimensionality of the problem from (\ref{eq:7}).
 For the multidimensional generalized global search algorithm (GSA) \cite{c5,c24}, developed in the framework of the information-statistical approach, the characteristic has the form
@@
 Again, we note that the above computational scheme is quite general.
-Within the framework of this scheme, many global search algorithms can be presented, as evidenced, in particular, by the examples given in (\ref{eq:16})-(\ref{eq:18}), as well as other characteristically represented algorithms (see, for example, \cite{c22,c23,c24,c25,c26,c27,c28,c29,c30,c31}).
+Within the framework of this scheme, many global search algorithms can be presented, as evidenced, in particular, by the examples given in (\ref{eq:16})--(\ref{eq:18}), as well as other characteristically represented algorithms (see, for example, \cite{c22,c23,c24,c25,c26,c27,c28,c29,c30,c31}).
 Thus, the parallelization methods discussed in Section \ref{sec:4} can be used for a sufficiently wide variety of global search algorithms.
@@
 
+\biboptions{sort&compress}
 \bibliography{mybibfile}

### 2020-05-31 2020/RSD_GPU/RSD2020.tex

@@
 \section{Multidimensional parallel global search algorithm}\label{Sec_GSA}
-Consider the problem of finding the global minimum of the \(N\)-dimensional function \(\varphi(y)\) in the hyperinterval  \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
+Let's consider the problem of finding the global minimum of the \(N\)-dimensional function \(\varphi(y)\) in the hyperinterval  \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
 We will assume that the function satisfies the Lipschitz condition with an a priori unknown constant \(L\).
@@
 Let \(n>1\) iterations of the method be performed, during which trials were carried out at \(k = k(n)\) points \(x^i, 1\leqslant i\leqslant k\).
-Then the trial points  \(x^{k+1},\dotsc,x^{k+p}\) of the next \((n+1)\)-th iteration are determined in accordance with the following rules.
+Then the trial points  \(x^{k+1},\dotsc,x^{k+p}\) of the next \((n+1)\)th iteration are determined in accordance with the following rules.
 
@@
 
-Step 4. The characteristics \(R(i),1\leqslant i\leqslant k+1\), arrange in descending order 
+Step 4. Arrange characteristics \(R(i),1\leqslant i\leqslant k+1\), in descending order 
 \begin{displaymath}
@@
 \begin{displaymath}
-x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},\; t_j=1, \; t_j=k+1
+x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},\; t_j=1, \; t_j=k+1,
 \end{displaymath}
@@
 As a result, when using evolvents, the images \(y' , y''\) that are close in \(N\)-dimensional space can correspond to rather distant preimages \(x' , x''\) on the interval \([0,1]\).
-This property leads to redundant calculations, because several limit points \(x' , x''\) of the test sequence generated by the method on the segment \([0,1]\), can correspond to a single limit point \(y\) in \(N\)-dimensional space.
+This property leads to redundant calculations, because several limit points \(x' , x''\) of the sequence of the trial points generated by the method on the segment \([0,1]\), can correspond to a single limit point \(y\) in \(N\)-dimensional space.
 
-A possible way to overcome this disadvantage is to use a set of evolvents (multiple mappings)
+A possible way to overcome this disadvantage is to use a set of evolvents (\textit{multiple mappings})
 \begin{displaymath}
@@
 \end{displaymath}
-And so on, until the calculation   according to (\ref{phiN}).
+And so on, until the calculation of $\varphi_N$  according to (\ref{phiN}).
 
@@
 
-Using new variables, the main relation of the multistep scheme (\ref{nestedScheme}) can be rewritten as
+Using new variables, the main relation of the nested scheme (\ref{nestedScheme}) can be rewritten as
 \begin{equation}
@@
 \end{equation}
-which are multidimensional, and to solve them, a method of reducing dimensionality based on Peano curves can be applied.
+which are multidimensional, and to solve them a method of reducing dimensionality based on Peano curves can be applied.
 
@@
 
-The solution to the resulting set of subproblems (\ref{subTasks}) can be organized in various ways.
+Solving the resulting set of subproblems (\ref{subTasks}) can be organized in various ways.
 The obvious method (elaborated in detail in \cite{Grishagin2015} for the nested optimization scheme and in \cite{Barkalov2014} for the block nested optimization scheme) is based on solving subproblems in accordance with the recursive order of their generation.
-However, there is a significant loss of information about the target function.
+However, in this case there is a significant loss of information about the target function.
 
@@
 	\begin{itemize}
-		\item to calculate the value of the function of the \(i\)-th level from (\ref{subTasks}), a new \(i+1\)-th level problem is generated in which trials are carried out, after which a new generated problem is included in the set of existing problems to be solved;
-		\item the iteration of the global search consists in choosing \(p\) (the most promising) problem from the set of existing problems in which trials are carried out; points for new trials are determined in accordance with the parallel global search algorithm from Section \ref{Sec_GSA};
+		\item to calculate the value of the function of the \(i\)-th level from (\ref{subTasks}), a new \((i+1)\)th level problem is generated in which trials are carried out, after which a new generated problem is included in the set of existing problems to be solved;
+		\item the iteration of the global search consists in choosing \(p\) (the most promising) problems from the set of existing problems in which trials are carried out; points for new trials are determined in accordance with the parallel global search algorithm from Section \ref{Sec_GSA};
 		\item the minimum values of functions from (\ref{subTasks}) are their current estimates based on the accumulated search information.
@@
 Therefore, at each iteration, subproblems with the maximum characteristic are selected for conducting the next trial.
-The trial either leads to the calculation of the value of the objective function \(\varphi_1(y)\) (if the selected subproblem belonged to the level \(j=M\)), or generates new subproblems according to (\ref{subTasks}) for \(j \geqslant  M-1\).
+The trial either leads to the calculation of the value of the objective function \(\varphi_1(y)\) (if the selected subproblem belonged to the level \(j=M\)), or generates new subproblems according to (\ref{subTasks}) for \(j \leqslant  M-1\).
 In the latter case, the newly generated problems are added to the current set of problems, their characteristics are calculated, and the process is repeated.
@@
 Data transfers from CPU to GPU will be minimal: one need only transfer the coordinates of the trial points to GPU, and get back the function values at these points.
-Functions that determine the processing of trial results in accordance with the algorithm, and require work with large amounts of accumulated search information, can be efficiently implemented on CPU.
+Functions that determine the processing of trial results in accordance with the algorithm, and require work with large amount of accumulated search information, can be efficiently implemented on CPU.
 
@@
 To organize parallel calculations, we will use a set of evolvents and a block adaptive dimensionality reduction scheme.
-We take a small number of nesting levels, 2 to 3, in which the original large dimensionality problem is divided into 2-3 nested lower dimensional subproblems.
-We will use multiple evolvents only at the upper level of nesting, corresponding to the u1 variable block.
+We take a small number of nesting levels, in which the original large dimensionality problem is divided into 2-3 nested lower dimensional subproblems.
+We will use multiple evolvents only at the upper level of nesting, corresponding to the variable $u_1$.
 This subproblem will be reduced to a set of one-dimensional problems that will be solved in parallel, each in a separate process.
@@
 In accordance with this scheme, nested subproblems  ${{\varphi }_{i}}({{u}_{1}},...,{{u}_{i}})=\underset{{{u}_{i+1}}\in {{D}_{i+1}}}{\mathop{\min }}\,{{\varphi }_{i+1}}({{u}_{1}},...,{{u}_{i}},{{u}_{i+1}})$  with \(i=1,\ldots ,M-2\) are solved using CPU only.
-The values of the optimized function are not calculated directly in these subproblems: the calculation the function value  ${{\varphi }_{i}}({{u}_{1}},...,{{u}_{i}})$ is a solution to the minimization problem at the next level.
-
-The subproblem of the last \((M - 1)\)-th level 
+The values of the function are not calculated directly in these subproblems: the calculation the function value  ${{\varphi }_{i}}({{u}_{1}},...,{{u}_{i}})$ is a solution to the minimization problem at the next level.
+The subproblem of the last \((M-1)\)-th level 
 \[
@@
  differs from all the previous subproblems; it calculates the values of the objective function, since ${{\varphi }_{M}}({{u}_{1}},...,{{u}_{M}})=\varphi ({{y}_{1}},...,{{y}_{N}})$.
-This subproblem transfers between CPU and GPU.
+This subproblem transfers data between CPU and GPU.
 
@@
 Each supercomputer node included two quad-core processors Intel Xeon X5570, two NVIDIA Tesla X2070 and 12 Gb RAM.
-When building the Globalizer system for running on the Lomonosov supercomputer, the GCC 4.3.0 compiler, CUDA 6.5 and Intel MPI 2017 were used.
+To build the program for running on the Lomonosov supercomputer, the GCC 4.3.0 compiler, CUDA 6.5 and Intel MPI 2017 were used.
 
@@
 Therefore, in order to simulate the computational complexity inherent in applied optimization problems \cite{Modorskii2016}, the calculation of the objective function in all experiments was complicated by additional calculations that do not change the form of the function and the location of its minima (summing the segment of the Taylor series).
-In the experiments carried out the average time for calculating the function values was 0.01 seconds, which exceeds the latency of the network or the data transfer time to the GPU.
+In the experiments carried out the average time for calculating the function values was 0.01 seconds, which exceeds the latency of the network or the data transfer time between CPU and GPU.
 
@@
 
-Below are the results of a numerical comparison of three sequential algorithms -- DIRECT \cite{Jones2009}, DIRECT\textit{l} \cite{Gablonsky2001} and Global Search Algorithm (GSA) from Section 2.
+Below are the results of a numerical comparison of three sequential algorithms: DIRECT \cite{Jones2009}, DIRECT\textit{l} \cite{Gablonsky2001} and Global Search Algorithm (GSA) from Section 2.
 A numerical comparison was carried out on the Simple and Hard function classes of dimension 4 and 5 from \cite{Gaviano2003}.
@@
 $\left\| {{y}^{k}}-{{y}^{*}} \right\| \leqslant \delta $.
-The size of the neighborhood was chosen (in accordance with \cite{Sergeyev2006}) as $\delta =\left\| b-a \right\|\sqrt[N]{\Delta }$, here $N$ is the dimension of the problem to be solved , $a$ and $b$ are the boundaries  of the search domain $D$, the parameter  $\Delta ={{10}^{-6}}$ for $N=4$ and $\Delta ={{10}^{-7}}$ for $N=5$.
+The size of the neighborhood was chosen (in accordance with \cite{Sergeyev2006}) as $\delta =\left\| b-a \right\|\sqrt[N]{\Delta }$, here $N$ is the dimension of the problem to be solved, $a$ and $b$ are the boundaries  of the search domain $D$, the parameter  $\Delta ={{10}^{-6}}$ for $N=4$ and $\Delta ={{10}^{-7}}$ for $N=5$.
 When using the GSA method for the Simple class, the parameter $r=4.5$ was selected, for the Hard class $r=5.6$; the parameter for constructing the Peano curve was $m=10$.

### 2020-06-05 2020/OPTIMA/optima_paper.tex
Optima paper

--- 2020-03-30 2020/PAVT_Sovrasov/vm_journal/vm_paper.tex
@@
-\documentclass{cmi}
-\usepackage{upgreek}
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.20 of 2017/10/04
+%
+\documentclass[runningheads]{llncs}
+%
+\usepackage{graphicx}
 
-% custom packages and commands
-\usepackage{doi}
+%custom packages
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{cite}
+\usepackage{subfig}
+\usepackage[labelfont=bf,labelsep=period]{caption}
+
+\usepackage[utf8]{inputenc}
+\inputencoding{utf8}
+\usepackage[english, russian]{babel}
 
@@
 \DeclareMathOperator{\sign}{sign}
-\renewcommand{\doitext}{DOI: }
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following line
+% to display URLs in blue roman font according to Springer's eBook style:
+% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
 \begin{document}
-
-\classify{УДК 004.021, 519.6}
-
-\renewcommand*{\thefootnote}{\fnsymbol{footnote}} %to make a number in end of title delete this line
-
-\title{ПАРАЛЛЕЛЬНЫЙ АЛГОРИТМ ДЛЯ ПОЛУЧЕНИЯ РАВНОМЕРНОГО ПРИБЛИЖЕНИЯ
-РЕШЕНИЙ МНОЖЕСТВА ЗАДАЧ ГЛОБАЛЬНОЙ ОПТИМИЗАЦИИ С~НЕЛИНЕЙНЫМИ ОГРАНИЧЕНИЯМИ%
-\footnote{Статья рекомендована к публикации программным комитетом Международной
-научной конференции <<Параллельные вычислительные технологии (ПаВТ) 2020>>.}%
+%
+\title{Parallel global optimization algorithm for obtaining uniform convergence when simultaneously solving a set of global optimization problems\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}}
+%
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
+%
+\author{Vladislav Sovrasov\and
+Konstantin Barkalov
 }
-
-\author{\copyright~2020 В.В.~Соврасов, К.A.~Баркалов}
-
-%\author{User1 \\ \href{mailto:me@somewhere.com}{me@somewhere.com}
-%	\and User2 \\ email %\href{mailto:someone@somewhere.com}{someone@somewhere.com} }
-
-\address{Нижегородский государственный университет им. Н.И.~Лобачевского\\
-  (603022 Нижний Новгород, пр.~Гагарина, д.~23, к.~2)}
-
-\email{E-mail: \href{mailto:sovrasov.vladislav@itmm.unn.ru}{sovrasov.vladislav@itmm.unn.ru},
-\href{mailto:konstantin.barkalov@itmm.unn.ru}{konstantin.barkalov@itmm.unn.ru}}
-
-\received{Поступила в редакцию: 27.02.2020}
-
-\maketitle{}
-
-\renewcommand*{\thefootnote}{\arabic{footnote}}
-\setcounter{footnote}{0}
-
-% оформление аннотации
-\begin{abstract}%
+%
+\authorrunning{V. Sovrasov et al.}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
+\email{sovrasov.vladislav@itmm.unn.ru},
+\email{konstantin.barkalov@itmm.unn.ru}
+}
+%
+\maketitle              % typeset the header of the contribution
+%
+\begin{abstract}
     В данной работе рассматривается построение параллельной версии алгоритма глобальной оптимизации, решающего одновременно множество задач с нелинейными ограничениями и получающего при этом равномерные оценки решений на этом множестве.
@@
 
-\keywords{глобальная оптимизация, параллельные вычисления, алгоритмы прямой оптимизации, равномерная сходимость}
+\keywords{Global optimization \and Uniform convergence \and Derivative-free optimization \and Parallel computations.}
 \end{abstract}
-\flushleft{\textbf{ОБРАЗЕЦ ЦИТИРОВАНИЯ}}
-\justifying
-\begin{citationplace}
-Соврасов~В.В., Баркалов~К.А. Параллельный алгоритм для получения равномерного приближения
-решений множества задач глобальной оптимизации с нелинейными ограничениями //
-Вестник ЮУрГУ. Серия: Вычислительная математика и информатика. 2020. Т.~X, №~Y. С.~Z1--Z2.
-DOI:~10.14529/cmseXXXXXX.
-\end{citationplace}
+%
+%
+%
 
-%Верхние колонтитулы
-\markboth{В.В.~Соврасов, К.A.~Баркалов}{Параллельный алгоритм для получения равномерного приближения\ldots}
-% Заголовки разделов формируются при помощи команд  \section{}, \subsection{}, \subsubsection{}
-\section*{Введение}
+\section{Введение}
 \label{sec-intro}
@@
 был сгенерирован смешанный класс, состоящий из 50 задач с двухмерными функциями GKLS и 50 задач с функциями \(F_{GR}\).
-На \figref{pic/2.png} и \figref{pic/4.png} представлены примеры линий уровня рассматриваемых задач. Допустимая область закрашена.
+На рис. \ref{fig:isolines} представлены примеры линий уровня рассматриваемых задач. Допустимая область закрашена.
 
-\fig{scale=0.3}{pic/2.png}{Пример линий уровня задачи с решением внутри допустимой области}
-\fig{scale=0.3}{pic/4.png}{Пример линий уровня задачи с решением на границе допустимой области}
+\begin{figure}[ht]
+    \centering
+    \subfloat[Решение внутри допустимой области]{
+    {\includegraphics[width=.5\textwidth]{pic/2.png}}}
+    \subfloat[Решение на границе допустимой области]{
+    {\includegraphics[width=.5\textwidth]{pic/4.png}}}
+    \caption{Линии уровня и точки испытаний ИАГП в двух синтетических задачах}
+    \label{fig:isolines}
+\end{figure}
 
@@
 Результаты решения тестовых задач последовательной и параллельной версией модифицированного ИАГП
-для решения множества задач представлены в \tabref{tab:speedup}. Для всех двухмерных классов задач параметр \(r=4.7\).
+для решения множества задач представлены в \ref{tab:speedup}. Для всех двухмерных классов задач параметр \(r=4.7\).
 В случае трехмерных задач \(r=4.7,\: \varepsilon_\nu=0.1\).
@@
 
-\tab{tab:speedup}{Результаты экспериментов на наборах синтетических задач}{
+\begin{table}
+  \centering
+  \caption{Результаты экспериментов на наборах синтетических задач}
+  \label{tab:speedup}
   \begin{tabular}{c|c|cccc}
@@
   \end{tabular}
-}
+\end{table}
 
 Для того, чтобы показать равномерную сходимость все тестовые задачи были также решены ИАГП
-в режиме решения отдельных задач. На \figref{pic/mixed_2d_max.pdf} и \figref{pic/mixed_2d_avg.pdf} указаны графики величин
+в режиме решения отдельных задач. На рис. \ref{fig:devs_mixed} указаны графики величин
 средних и максимальных расстояний от реальных оптимумов до их текущих оценок при решении
@@
 
-\fig{scale=0.5}{pic/mixed_2d_max.pdf}{Динамика величины \(D_{max}\) в процессе решения множества двухмерных задач,
+\begin{figure}[ht]
+    \centering
+    \subfloat[\(D_{max}\)]{
+    \label{fig:max_dev} {\includegraphics[width=.5\textwidth]{pic/mixed_2d_max.pdf}}}
+    \subfloat[\(D_{avg}\)]{
+    \label{fig:avg_dev} {\includegraphics[width=.5\textwidth]{pic/mixed_2d_avg.pdf}}}
+    \caption{Динамика величин \(D_{avg}\) и \(D_{max}\) в процессе решения множества двухмерных задач,
     порождённых двумя разными генераторами GKLS и \(F_{GR}\)}
-\fig{scale=0.5}{pic/mixed_2d_avg.pdf}{Динамика величины \(D_{avg}\) в процессе решения множества двухмерных задач,
-порождённых двумя разными генераторами GKLS и \(F_{GR}\)}
+    \label{fig:devs_mixed}
+\end{figure}
 
@@
 
-\fig{scale=0.5}{pic/single_mco.pdf}{Численная оценки множества Парето в задаче (\ref{eq:mco_probem}),
-                                    полученная при раздельном решении задач}
-\fig{scale=0.5}{pic/multi_mco.pdf}{Численная оценки множества Парето в задаче (\ref{eq:mco_probem}),
-                                   полученная при совместном решении задач}
+\begin{figure}[ht]
+    \centering
+    \subfloat[ИАГП, раздельное решение задач]{
+    \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{pic/single_mco.pdf}}}
+    \subfloat[ИАГП для множества задач]{
+    \label{fig:mco_pareto_2} {\includegraphics[width=.5\textwidth]{pic/multi_mco.pdf}}}
+    \caption{Численные оценки множества Парето в задаче (\ref{eq:mco_probem}), полученные после 2500 испытаний}
+    \label{fig:mco_pareto}
+\end{figure}
 
-На \figref{pic/single_mco.pdf} и \figref{pic/multi_mco.pdf}
+На рис. \ref{fig:mco_pareto_1} и рис. \ref{fig:mco_pareto_2}
 представлены графики решений, полученных каждым из методов.
 Все графики качественно совпадают с указанным в \cite{BinhKorn1999} (авторы не предоставили другой
-информации и решениях для сравнения). Видно, что на \figref{pic/single_mco.pdf}
+информации и решениях для сравнения). Видно, что на рис. \ref{fig:mco_pareto_1}
 кривая Парето имеет вогнутости, что не соответствует решению, указанному в \cite{BinhKorn1999} и
@@
 \begin{displaymath}
-  SP(S)=\sqrt{\frac{1}{|S|-1} \sum_{i=1}^{|S|} (\overline{d}-d_i)},
+  SP(S)=\sqrt{\frac{1}{|S|-1} \sum_{i=1}^{|S|} (\overline{d}-d_i)^2},
   \:\overline{d}=mean\{d_i\},\:d_i=\min_{s_i,s_j\in S:s_i\ne s_j}||F(s_i)-F(s_j)||_1,\: F=(f_1,f_2)
@@
 
-\section*{Заключение}
+\section{Заключение}
 
@@
 
-\vspace{1em}
-{\it Исследование выполнено при поддержке РНФ, проект №\,16-11-10150.}
+%
+% ---- Bibliography ----
+%
+% BibTeX users should specify bibliography style 'splncs04'.
+% References will then be sorted and formatted in the correct style.
+%
+% \bibliographystyle{splncs04}
+% \bibliography{mybibliography}
+%
+\begin{thebibliography}{8}
+\bibitem{ref_article1}
+Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
 
-\begin{biblio}
-  \bibitem{BarkalovLebedev2017}
-  Barkalov K., Lebedev I.
-  \newblock Comparing two approaches for solving constrained global optimization
-    problems
-  \newblock // Learning and Intelligent Optimization. Springer International
-    Publishing, Cham. 2017. P. 301--306.
-  \newblock \doi{10.1007/978-3-319-69404-7_22}
+\bibitem{ref_lncs1}
+Author, F., Author, S.: Title of a proceedings paper. In: Editor,
+F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
+Springer, Heidelberg (2016). \doi{10.10007/1234567890}
 
-  \bibitem{BarkalovLebedev2017_2}
-  Barkalov K., Lebedev I.
-  \newblock Parallel algorithm for solving constrained global optimization
-    problems
-  \newblock // Parallel Computing Technologies. Springer International
-    Publishing, Cham. 2017. P. 396--404.
-  \newblock \doi{10.1007/978-3-319-62932-2_38}
+\bibitem{ref_book1}
+Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
+Location (1999)
 
-  \bibitem{BarkalovStrongin2018}
-  Barkalov K., Strongin R.
-  \newblock Solving a set of global optimization problems by the parallel
-    technique with uniform convergence //
-  \newblock Journal of Global Optimization. 2018. Vol. 71, no. 1. P. 21--36.
-  \newblock \doi{10.1007/s10898-017-0555-4}
+\bibitem{ref_proc1}
+Author, A.-B.: Contribution title. In: 9th International Proceedings
+on Proceedings, pp. 1--2. Publisher, Location (2010)
 
-  \bibitem{Beiranvand2017}
-  Beiranvand V., Hare W., Lucet Y.
-  \newblock Best practices for comparing optimization algorithms //
-  \newblock Optimization and Engineering. 2017. Vol. 18, no. 4. P. 815--848.
-  \newblock \doi{10.1007/s11081-017-9366-1}
-
-  \bibitem{Ehrgott2005}
-  Ehrgott M.
-  \newblock Multicriteria Optimization.
-  \newblock Springer-Verlag, Berlin, Heidelberg. 2005.
-  \newblock \doi{10.1007/3-540-27659-9}
-
-  \bibitem{Evtushenko2013}
-  Evtushenko Y., Posypkin M.
-  \newblock A deterministic approach to global box-constrained optimization //
-  \newblock Optimization Letters. 2013. Vol. 7 P. 819--829.
-  \newblock \doi{10.1007/s11590-012-0452-1}
-
-  \bibitem{Gaviano2003}
-  {Gaviano M., Kvasov D.E., Lera D., Sergeev Ya.D.}
-  \newblock Software for generation of classes of test functions with known local
-    and global minima for global optimization //
-  \newblock ACM Transactions on Mathematical Software. 2003. Vol. 29, no. 4. P.
-    469--480.
-  \newblock \doi{10.1145/962437.962444}
-
-  \bibitem{GergelBarkalov2019}
-  Gergel V., Barkalov K., Lebedev I., Rachinskaya M., Sysoyev A.
-  \newblock A flexible generator of constrained global optimization test
-    problems // Proceedings LeGO - 14th International Global Optimization Workshop. 8–21 September 2018, Leiden, The Netherlands. Vol. 2070, no. 1. P. 020009.
-  \newblock \doi{10.1063/1.5089976}
-
-  \bibitem{Jones2009}
-  {Jones D.R.}
-  \newblock The direct global optimization algorithm
-  \newblock // The Encyclopedia of Optimization. Springer, Heidelberg. 2009. P.
-    725--735.
-  \newblock \doi{10.1007/978-0-387-74759-0_128}
-
-  \bibitem{Paulavicius2011}
-  {Paulavivcius R., Zilinskas J., Grothey A.}
-  \newblock Parallel branch and bound for global optimization with combination of
-    lipschitz bounds //
-  \newblock Optimization Methods and Software. 1997. Vol. 26, no. 3. P. 487--498.
-  \newblock \doi{10.1080/10556788.2010.551537}
-
-  \bibitem{RiquelmeLucken2015}
-  {Riquelme} N., {Von Lucken} C., {Baran} B.
-  \newblock Performance metrics in multi-objective optimization
-  \newblock // 2015 Latin American Computing Conference (CLEI). 2015. P. 1--11.
-  \newblock \doi{10.1109/clei.2015.7360024}
-
-  \bibitem{SergeyevKvasov2017}
-  Sergeyev Y., Kvasov D. Deterministic Global Optimization. 2017. 136 p.
-  \newblock \doi{10.1007/978-1-4939-7199-2}
-
-  \bibitem{Sergeyev2013}
-  {Sergeyev Y.D., Strongin R.G., Lera D.} Introduction to global optimization
-    exploiting space-filling curves.
-  \newblock Springer Briefs in Optimization, Springer, New York. 2013.
-  \newblock \doi{10.1007/978-1-4614-8042-6}
-
-  \bibitem{Strongin2000}
-  {Strongin R.G., Sergeyev Ya.D.} Global optimization with non-convex
-    constraints. Sequential and parallel algorithms.
-  \newblock Kluwer Academic Publishers, Dordrecht. 2000. 704 p.
-  \newblock \doi{10.1007/978-1-4615-4677-1}
-
-  \bibitem{BinhKorn1999}
-  To T.B., Korn B.
-  \newblock MOBES: A multiobjective evolution strategy for constrained optimization problems
-  // Proceedings of the third international conference on genetic algorithms (Mendel 97). 1997. P. 176--182.
-
-  \bibitem{grishaginClass}
-  {Гришагин В.А.}
-  \newblock {Операционные характеристики
-    некоторых алгоритмов глобального поиска}
-  \newblock // Проблемы статистической
-    оптимизации. Зинатне, Рига. 1978. №7. С. 198–206.
-
-  \bibitem{Norkin1992}
-  {Норкин В. И.}
-  \newblock О методе Пиявского для решения общей
-    задачи глобальной оптимизации //
-  \newblock Журнал вычислительной математики и математической физики. 1992. № 32 С. 992--1006.
-\end{biblio}
-
-% Сведения об авторах (Ф.И.О.(полностью), степень, звание, кафедра (отдел), вуз, e-mail)
-Владислав Валерьевич Соврасов, аспирант, Кафедра математического обеспечения и
-суперкомпьютерных технологий, Нижегородский государственный университет им. Н.И.~Лобачевского (Нижний Новгород, Российская Федерация)
-
-Константин александрович Баркалов, к.ф.м.н., доцент, Кафедра математического обеспечения и
-суперкомпьютерных технологий, Нижегородский государственный университет им. Н.И.~Лобачевского (Нижний Новгород, Российская Федерация)
-
-\newpage
-\classify{} % Код MSC не нужен
-%\classify{MSC 00000} % Коды MSC см. http://www.ams.org/msc/
-
-\title{\uppercase{Parallel global optimization algorithm for~obtaining uniform convergence
-when~simultaneously solving a~set of~global optimization problems}}
-
-\author{\copyright~2020 V.V.~Sovrasov, K.A.~Barkalov}
-
-\address{Lobachevsky State University of Nizhni Novgorod, Russia (pr. Gagarina 23(2), Nizhni Novgorod, 603022 Russia)}
-
-\email{E-mail: \href{mailto:sovrasov.vladislav@itmm.unn.ru}{sovrasov.vladislav@itmm.unn.ru},
-\href{mailto:konstantin.barkalov@itmm.unn.ru}{konstantin.barkalov@itmm.unn.ru}}
-
-\received{Received: 27.02.2020}
-
-\maketitle{}
-
-\begin{abstract}%
-
-In this work building of a parallel version of a method simultaneously solving a set of constrained global optimization problems is considered.
-This method converges uniformly to solutions of all the problems.
-That allows the method to arrange computational resources in an optimal way, since uniform convergence guarantees approximately equal precision of numerical solutions at the whole set of problems at the each iteration of optimization.
-The algorithm assigns a priority to each problem, and then at the each iteration carries out calculation of objective functions and constraints in several problems in parallel.
-If the method stops at any arbitrary moment, in all the problems numerical sulutions with the similar accuracy will be obtained.
-Sets of similar global optimization problems appear for an instance after scalarization of multi-objective problems or when a global optimization problem has a discrete parameter which takes a finite number of possible values.
-The considered method uses Peano-type curves to transform multidimensional problems into univariate ones.
-Efficiency of the implemented parallel algorithm is evaluated on several sets of synthetically generated constrained global optimization problems and on a scalarized multi-objective problem.
-Also the uniform convergence was confirmed numerically by validation quality of intermediate solutions during the optimization process.
-
-  \keywordsen{global optimization, parallel computations, derivative-free optimization, uniform convergence}
-\end{abstract}
-
-\begin{flushleft}
-  \textbf{FOR CITATION}
-\end{flushleft}
-\justifying
-
-\begin{citationplace}
-Sovrasov~V.V., Barkalov~K.A. Parallel global optimization algorithm for obtaining uniform convergence
-when simultaneously solving a set of global optimization problems.
-Bulletin of the South Ural State University.
-Series: Computational Mathematics and Software Engineering.
-2020.
-Vol.~X, no.~Y.
-P.~Z1--Z2.
-(in Russian) DOI: 10.14529/cmseXXXXXX.
-\end{citationplace}
-
-\begin{biblio_lat}
-%\itemsep -2pt
-
-\bibitem{BarkalovLebedev2017}
-Barkalov K., Lebedev I.
-\newblock Comparing two approaches for solving constrained global optimization
-  problems.
-\newblock Learning and Intelligent Optimization. Springer International
-  Publishing, Cham. 2017. P. 301--306.
-\newblock \doi{10.1007/978-3-319-69404-7_22}
-
-\bibitem{BarkalovLebedev2017_2}
-Barkalov K., Lebedev I.
-\newblock Parallel algorithm for solving constrained global optimization
-  problems.
-\newblock Parallel Computing Technologies. Springer International
-  Publishing, Cham. 2017. P. 396--404.
-\newblock \doi{10.1007/978-3-319-62932-2_38}
-
-\bibitem{BarkalovStrongin2018}
-Barkalov K., Strongin R.
-\newblock Solving a set of global optimization problems by the parallel
-  technique with uniform convergence.
-\newblock Journal of Global Optimization. 2018. Vol. 71, No. 1. P. 21--36.
-\newblock \doi{10.1007/s10898-017-0555-4}
-
-\bibitem{Beiranvand2017}
-Beiranvand V., Hare W., Lucet Y.
-\newblock Best practices for comparing optimization algorithms.
-\newblock Optimization and Engineering. 2017. Vol. 18, No. 4. P. 815--848.
-\newblock \doi{10.1007/s11081-017-9366-1}
-
-\bibitem{Ehrgott2005}
-Ehrgott M.
-\newblock Multicriteria Optimization.
-\newblock Springer-Verlag, Berlin, Heidelberg. 2005.
-\newblock \doi{10.1007/3-540-27659-9}
-
-\bibitem{Evtushenko2013}
-Evtushenko Y., Posypkin M.
-\newblock A deterministic approach to global box-constrained optimization.
-\newblock Optimization Letters 2013. Vol. 7 P. 819--829.
-\newblock \doi{10.1007/s11590-012-0452-1}
-
-\bibitem{Gaviano2003}
-{Gaviano M., Kvasov D.E., Lera D., Sergeev Ya.D.}
-\newblock Software for generation of classes of test functions with known local
-  and global minima for global optimization.
-\newblock ACM Transactions on Mathematical Software. 2003. Vol. 29, No. 4. P.
-  469--480.
-\newblock \doi{10.1145/962437.962444}
-
-\bibitem{GergelBarkalov2019}
-Gergel V., Barkalov K., Lebedev I., Rachinskaya M., Sysoyev A.
-\newblock A flexible generator of constrained global optimization test
-  problems // Proceedings LeGO - 14th International Global Optimization Workshop. 8–21 September 2018, Leiden, The Netherlands. Vol. 2070, no. 1. P. 020009.
-\newblock \doi{10.1063/1.5089976}
-
-\bibitem{Jones2009}
-{Jones D.R.}
-\newblock The direct global optimization algorithm.
-\newblock The Encyclopedia of Optimization. Springer, Heidelberg. 2009. P.
-  725--735.
-\newblock \doi{10.1007/978-0-387-74759-0_128}
-
-\bibitem{Paulavicius2011}
-{Paulavivcius R., Zilinskas J., Grothey A.}
-\newblock Parallel branch and bound for global optimization with combination of
-  lipschitz bounds.
-\newblock Optimization Methods and Software. 1997. Vol. 26, No. 3. P. 487--498.
-\newblock \doi{10.1080/10556788.2010.551537}
-
-\bibitem{RiquelmeLucken2015}
-{Riquelme} N., {Von Lucken} C., {Baran} B.
-\newblock Performance metrics in multi-objective optimization.
-\newblock2015 Latin American Computing Conference (CLEI). 2015. P. 1--11.
-\newblock \doi{10.1109/clei.2015.7360024}
-
-\bibitem{SergeyevKvasov2017}
-Sergeyev Y., Kvasov D. Deterministic Global Optimization. 2017. 136 p.
-\newblock \doi{10.1007/978-1-4939-7199-2}
-
-\bibitem{Sergeyev2013}
-{Sergeyev Y.D., Strongin R.G., Lera D.} Introduction to global optimization
-  exploiting space-filling curves.
-\newblock Springer Briefs in Optimization, Springer, New York. 2013.
-\newblock \doi{10.1007/978-1-4614-8042-6}
-
-\bibitem{Strongin2000}
-{Strongin R.G., Sergeyev Ya.D.} Global optimization with non-convex
-  constraints. Sequential and parallel algorithms.
-\newblock Kluwer Academic Publishers, Dordrecht. 2000. 704 p.
-\newblock \doi{10.1007/978-1-4615-4677-1}
-
-\bibitem{BinhKorn1999}
-To T.B., Korn B.
-\newblock MOBES: A multiobjective evolution strategy for constrained optimization problems //
-Proceedings of the third international conference on genetic algorithms (Mendel 97). 1997. P. 176--182.
-
-\bibitem{grishaginClass}
-{Grishagin V.}
-\newblock {Operating characteristics of some global search algorithms.}
-\newblock Problems of Stochastic Search. Zinatne, Riga. 1978. No. 7. P. 198–206. (in Russian)
-
-\bibitem{Norkin1992}
-{Norkin V. I.}
-\newblock Towards Pijavskyj's method for solving common global optimization problem.
-\newblock Computational Mathematics and Mathematical Physics. 1992. Vol. 32, No. 7. P. 992--1006. (in Russian)
-
-\end{biblio_lat}
-
+\bibitem{ref_url1}
+LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
+Oct 2017
+\end{thebibliography}
 \end{document}

### 2020-06-13 2016/JoGO_Adaptive/paper.tex

\RequirePackage{fix-cm}
\documentclass[smallextended]{svjour3}
\smartqed

\usepackage{bbm}
\usepackage{cite}
\usepackage{tikz}
\usepackage{float}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{changepage}
\usepackage[utf8]{inputenc}

% Command redefinitions
\let\origref\ref
\renewcommand{\ref}[1]{(\origref{#1})}

\journalname{Journal of Global Optimization}

\begin{document}
\title{Adaptive nested optimization scheme for multidimensional global search}

\author{Victor Gergel       \and
        Vladimir Grishagin  \and
        Alexander Gergel
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{Department of Software, Nizhni Novgorod State University, pr. Gagarina, 23, 603950, Nizhni Novgorod, Russia \\
           \\
           V. Gergel \at
              \email{gergel@unn.ru}
           \and
           V. Grishagin \at
              \email{vagris@unn.ru}
           \and
           A. Gergel \at
              \email{gergelm@unn.ru} \\
           \\
           Tel: +7-831-462-33-56
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
Methods for solving the multidimensional multiextremal optimization problems using the nested optimization scheme are considered. A novel approach for solving the multidimensional multiextremal problems based on the adaptive nested optimization has been proposed. This approach enables to develop methods of the global optimum search which are more efficient in comparison with the ones on the base of the traditional nested optimization scheme. The new approach provides advantages due to better usage of the information on the problem in the course of optimization. A general scheme of a adaptive nested optimization is described. A theoretical substantiation of the method convergence is given for the case when for solving the univariate subproblems within the nested scheme an information algorithm of global search is used. Results of numerical experiments on the well-known classes of the test multiextremal functions confirming the efficiency of the proposed scheme are presented.

\keywords{Multidimensional multiextremal optimization \and Global search algorithms \and Dimensionality reduction \and Adaptive nested optimization scheme}
\end{abstract}

\section{Introduction}
\label{intro}
In the present article, methods for solving multiextremal problems widespread in the real life applications are considered. The increasing complexity of mathematical models for more adequate description of objects, phenomena, and systems under study results in a significant increase of the complexity in analysis of  these models. Numerical experiments become the primary tool for such an analysis. The increasing complexity of the problems results in the necessity of a purposeful selection of the variants in the process of the optimal solution search. The essence of the purposeful selection consists in eliminating from further consideration many obviously unpromising cases on the base of the analysis of a small portion of the variants in order to concentrate further search in the subdomains containing the best variants. However, the development of the efficient optimization algorithms in the multidimensional case is labor-consuming often. A possible way in this situation consists in applying different schemes of dimensionality reduction for the development of the global search methods.

The multidimensional multiextremal (\textit{global}) optimization problem without constraints can be defined as a problem of finding the minimal value of a real function $\varphi(x)$
%
\begin{equation}
\label{eq:1}
\varphi(y^*) = \min \{ \varphi(y): y \in D \},
\end{equation}
%
and its global minimizer $y^* \in D$, where
%
\begin{equation}
\label{eq:2}
D = \{ y \in \mathbbm{R}^N : y_i \in [a_i, b_i],  1 \leq i \leq N \}
\end{equation}
%
is the search domain, which  is a hyperparallelepiped in the $N$-dimensional Euclidean space $\mathbbm{R}^N$.

The global optimization problems are the subject of the extensive research (see, for example, comprehensive bibliography in \cite{Ref4, Ref5, Ref14, Ref15, Ref16, Ref22, Ref23, Ref24, Ref33, Ref36, Ref37, Ref38, Ref39}). At the same time, these problems are very computation-consuming. Thus, if  for the solution search  the method of full scanning the nodes of a uniform grid is used with the accuracy $\varepsilon > 0$ for each coordinate, the total number of the grid nodes in the search domain can be estimated as
%
\begin{equation*}
K = K(\varepsilon) \approx \prod\limits_{j = 1}^N \big[ (b_j - a_j) / \varepsilon \big].
\end{equation*}
%
The decrease of the computational costs is possible only thanks to the highly efficient global search algorithms, which generate the dense grids in the vicinity of the sought minimum of the minimized function only. The design of such efficient algorithms directly for the multidimensional case is a rather complex problem. As a result, using for this goal various \textit{dimensionality reduction schemes} is one of the directions in the field of global optimization. Within the framework of this direction, the following three main approaches could be outlined:

The first dimensionality reduction method is based on the well-known fundamental property, according to which an $N$-hyperparallelepiped from~\ref{eq:2} and the interval of the real axis $[0, 1]$ are the sets of equal cardinality, and the interval $[0, 1]$ can be mapped onto the hyperparallelepiped $D$ unambiguously and continuously (see \cite{Ref1, Ref30, Ref36}). The mappings of this kind are usually referred to as Peano \textit{evolvents} or curves.

Assume $y(x), x \in [0, 1]$, to be a Peano curve and the function $\varphi(y)$ from~\ref{eq:1} to be a continuous one. Then, the optimization problem for a multidimensional function $\varphi(y)$ can be reduced to a problem of optimization of a one-dimensional reduced function $\varphi(y(x))$
%
\begin{equation*}
\varphi(y(x^*)) = \min \{ \varphi(y(x)): x \in [0, 1] \} = \min \{ \varphi(y): y \in D \}.
\end{equation*}
This approach has served a basis for the development of many efficient multidimensional global search algorithms (see \cite{Ref9, Ref11, Ref13, Ref30, Ref34, Ref36}).

In the framework of the second approach, the search domain is divided into the subdomains by means of a partition strategy, each subdomain is estimated from the point of view of its importance, or prospectivity for continuation of the search and the next iteration point is chosen in the subdomain of maximal importance. This approach is close to univariate scheme of characteristical algorithms \cite{Ref11} and the methods based on multidimensional partitioning are presented, for instance, in \cite{Ref17, Ref19, Ref20, Ref21, Ref22, Ref23, Ref24}.

Finally, the well-known approach based on the usage of the one-dimensional optimization methods for solving the multidimensional optimization problems  consists in the application of the \textit{nested optimization scheme}. According to this scheme, the solving of a multidimensional optimization problem could be obtained by means of solving  a sequence of the recursively nested one-dimensional problems (see, for example \cite{Ref2, Ref3, Ref8, Ref12, Ref25, Ref28, Ref31, Ref36}). The core of this approach is the relation
\begin{equation}
\label{eq:3}
\min_{y \in D} \varphi(y) = \min_{y_1 \in [a_1, b_1]} \dots \min_{y_N \in [a_N, b_N]} \varphi(y_1, \dots, y_N).
\end{equation}

According to~\ref{eq:3} the solving of a multidimensional multiextremal optimization problem is reduced to the solving of the one-dimensional problem:
%
\begin{equation}
\label{eq:4}
\varphi^* = \min_{y \in D} \varphi(y) = \min_{y_1 \in [a_1, b_1]} \widetilde{\varphi}_1(y_1),
\end{equation}
%
where
%
\begin{equation}
\label{eq:5}
\widetilde{\varphi}_i(y_i) = \varphi_i(y_1, \dots, y_i) = \min_{y_1 \in [a_{i + 1}, b_{i + 1}]} \varphi_{i + 1}(y_1, \dots, y_i, y_{i + 1}), 1 \leq i \leq N,
\end{equation}
%
\begin{equation}
\label{eq:6}
\varphi_N(y_1, \dots, y_N) = \varphi(y_1, \dots, y_N).
\end{equation}
%
The one-dimensional function $\widetilde{\varphi}_1(y_1)$ introduced in~\ref{eq:4} is constructed according to general recursive rule~\ref{eq:5}: in order to calculate the value $\widetilde{\varphi}_1(y_1)$ for given value of the variable $y_1 = \hat{y}_1$ it is necessary to perform the minimization of the function
\begin{equation*}
\widetilde{\varphi}_2(y_2) = \varphi_2(\hat{y}_1, y_2).
\end{equation*}
%
When performing this optimization, the function $\widetilde{\varphi}_2$ is a one-dimensional one as well since the value of the variable $y_1$ is predefined and fixed. Next, in turn, in order to calculate the value $\widetilde{\varphi}_2(y_2)$ at the point $y_2 = \hat{y}_2$ it is necessary to perform the minimization of the univariate function
\begin{equation*}
\widetilde{\varphi}_3(y_3) = \varphi_3(\hat{y}_1, \hat{y}_2, y_3),
\end{equation*}
%
etc.

The scheme~\ref{eq:4}--\ref{eq:6} in the case of  the Lipschitzian one-dimensional functions $\widetilde{\varphi}_i$ combined with the one-dimensional algorithms \cite{Ref7, Ref25, Ref32, Ref34, Ref35, Ref36} has served as the basis for designing  a number of the multidimensional methods \cite{Ref3, Ref12, Ref25, Ref26, Ref36}.

Concluding the characterization of the nested optimization scheme, it should be noted that this approach to the design of the multidimensional methods has also a number of disadvantages. When solving the particular subproblem~\ref{eq:5} the information on the behavior of the objective function obtained during solving other subproblems is not used in any way. This is the fundamental cause of the loss in the efficiency of the nested scheme. On the one hand, this makes the computational scheme simpler because allows to avoid storing and processing large amount of information. On the other hand, the refusal from the information worsens the search process planning quality.
Besides, when using this scheme, the precision in the termination condition providing the termination of any recursively nested one-dimensional problem like~\ref{eq:5} should be defined a priori. If the accuracy appears to be not enough, it is required to repeat the solving of the problem as a whole from the beginning with an increased precision. At the same time, the use of too high precision can result in an essential increase in the global search duration, and breaking the computations before the termination condition is satisfied (because of exceeding of the feasible computation time) in this case can lead to the case when the estimate of the sought solution is obtained in some subdomain of the search domain $D$ only.

The rest of the paper is organized as follows. Section~\origref{sec:2} describes a new generalization of the nested optimization scheme. Section~\origref{sec:3} is devoted to the theoretical substantiation of the new scheme concerning convergence properties. Section~\origref{sec:4} contains results of computational experiments. The last section gives a succinct summary of the paper.

\section{Adaptive nested optimization  scheme}
\label{sec:2}
Let us give a generalization of the basic scheme of dimensionality reduction~\ref{eq:3}--\ref{eq:6} allowing to overcome some disadvantages of the approach mentioned above. The essence of this generalization consists in the elimination of the principle of a strict subordination for the minimization problems of the one-dimensional functions $\widetilde{\varphi}_i(y_i), 1 \leq i \leq N$, generated within the framework of the nested optimization scheme. In the new approach all these problems are supposed to be solved simultaneously.

The preliminary description of the proposed generalization is as follows.

Let us introduce the following notation for brevity
\begin{equation}
\label{eq:7}
\nu_i = (y_1, \dots, y_i), 1 \leq i \leq N.
\end{equation}

Then, the one-dimensional functions $\widetilde{\varphi}_i(y_i), 1 < i \leq N$, generated within the framework of the nested optimization scheme can be rewrote in the form
\begin{equation}
\label{eq:8}
\widetilde{\varphi}_i(y_i) = \varphi_i(\nu_{i - 1}, y_i), 1 < i \leq N,
\end{equation}
%
where the vector $\nu_{i - 1}$ is a fixed one.

Let us describe in more details the procedure of computations for the initial (basic) scheme of dimensionality reduction.

At every global search iteration $i, 1 \leq i \leq k$, the optimization algorithm begins its computations from a one-dimensional function of the first level $\widetilde{\varphi}_1(y_1)$. The algorithm selects a current search point $y_1 = \hat{y}_1$ taking into account the search information obtained already. In order to get the function value $\widetilde{\varphi}_1(y_1)$ at this point, a new one-dimensional function $\varphi_2(\hat{y}_1, y_2)$ to be minimized is generated. To do so, the algorithm should suspend the minimization of the function $\widetilde{\varphi}_1(y_1)$ and start minimizing the function $\varphi_2(\hat{y}_1, y_2)$. In the optimization process of the function $\varphi_2(\hat{y}_1, y_2)$, the algorithm generates a sequence of the points
\begin{equation*}
\{ y_2^j, 1 \leq j \leq k \},
\end{equation*}
%
for each of which it is necessary to execute the minimization of the third level $\varphi_3(\hat{y}_1, \hat{y}_2, y_3)$ in the same strict order: when determining the next search point before the transition to the next search iteration, the solving of a one-dimensional problem of the next dimensionality reduction level should be carried out first. This process should be fulfilled recursively until the last $N^{\text{th}}$ level. The resulting hierarchical scheme of generating and solving the one-dimensional problems is presented in Fig.~\origref{fig:1}.

\begin{figure}[H]
\centering
\begin{tikzpicture}
\usetikzlibrary{arrows,positioning,calc}
\tikzset{
  task/.style = {draw, text width=27mm, text height=3mm, rectangle, rounded corners=2pt, align=center},
  arrow/.style ={->, >=latex'}
}

\node[task] (task11) {$\min\limits_{y_1 \in [a_1, b_1]}\varphi_1(y_1)$};

\node[task, below=5mm of task11] (task22) {$\min\limits_{y_2 \in [a_2, b_2]}\varphi_2(\hat{\nu}_1^2, y_2)$};
\node[task, left=2mm of  task22] (task21) {$\min\limits_{y_2 \in [a_2, b_2]}\varphi_2(\hat{\nu}_1^1, y_2)$};
\node[task, right=2mm of task22] (task23) {$\min\limits_{y_2 \in [a_2, b_2]}\varphi_2(\hat{\nu}_1^3, y_2)$};

\node[task, below=5mm of task22] (task32) {$\min\limits_{y_3 \in [a_3, b_3]}\varphi_3(\hat{\nu}_2^1, y_3)$};
\node[task, left=2mm of  task32] (task31) {$\min\limits_{y_3 \in [a_3, b_3]}\varphi_3(\hat{\nu}_2^2, y_3)$};
\node[task, right=2mm of task32] (task33) {$\min\limits_{y_3 \in [a_3, b_3]}\varphi_3(\hat{\nu}_2^3, y_3)$};

\draw[arrow] (task11.south) -- ++(0,-0.2) -|  (task21.north);
\draw[arrow] (task11.south) -- ++(0,-0.2) -|  (task22.north);
\draw[arrow] (task11.south) -- ++(0,-0.2) -|  (task23.north);

\draw[arrow] (task22.south) -- ++(0,-0.2) -|  (task31.north);
\draw[arrow] (task22.south) -- ++(0,-0.2) -|  (task32.north);
\draw[arrow] (task22.south) -- ++(0,-0.2) -|  (task33.north);

\node at ($ (task31.north) + (-0.5,  0.2) $) {$\dots$};
\node at ($ (task33.north) + ( 0.5,  0.2) $) {$\dots$};
\node at ($ (task32.south) + ( 0.0, -0.3) $) {$\dots$};
\node[right=0.5mm of task23] {$\dots$};
\node[right=0.5mm of task33] {$\dots$};

\end{tikzpicture}
\caption{\label{fig:1}The hierarchical scheme of generating and solving the one-dimensional problems in the nested optimization scheme}
\end{figure}

The functions generated during the optimization process have a strict hierarchical structure of subordination in the form of a tree. The structure of this tree changes dynamically in the course of solving the multidimensional task. The computation of the function value $\varphi_i(\hat{\nu}_{i - 1}, y_i)$ of the $i^{\text{th}}$ level, $1 \leq i < N$, at some point   requires solving all problems in one of the subtrees of the $(i + 1)^{\text{th}}$ level. The functions $\varphi_N(\hat{y}_{N - 1}, y_N)$ of the $N^{\text{th}}$ level are the leaves of the problem tree, their values are the values of the objective function $\varphi(y), y \in D$, and are computed directly without the nested optimization.

The \textit{proposed generalization of the basic scheme} of the dimensionality reduction consists in the elimination of the strict order of the subproblem solving according to their hierarchy in the problem tree, when the solving of the problem of the $i^{\text{th}}$ level, $1 \leq i < N$, does not require the complete  solving of the nested problems of the $(i + 1)^{\text{th}}$ level. Within the framework of the novel approach all generated problems are considered simultaneously and are equal in rights. The proposed approach will be hereinafter referred to as \textit{the adaptive nested optimization scheme}.

For description of the approach, let us renumber problems~\ref{eq:5} in the following way. Since the nested optimization scheme starts from the solving of the problem
\begin{equation}
\label{eq:9}
\min_{y_1 \in [a_1, b_1]} \varphi_1(y_1),
\end{equation}
%
let us assign to this problem the index $l = 1$. The solving of this problem begins with the computation of the function $\widetilde{\varphi}_1(y_1)$ at some point $\hat{y}_1$ that generates the problem
\begin{equation*}
\min_{y_2 \in [a_2, b_2]} \varphi_2(\hat{y}_1, y_2),
\end{equation*}
%
to which the index $l = 2$ is assigned, etc.

In general, when $Q \leq 1$ subproblems~\ref{eq:5} have been initiated already, the index $Q + 1$ is assigned to a newly generated subproblem, and the number of subproblems $Q$ is incremented by $1$.

Thus, at every moment of execution of the nested optimization scheme there is a set $F_Q$ consisting of the problems of the kind
\begin{equation}
\label{eq:10}
\min_{x \in [a, b]} f_l(x), 1 \leq l \leq Q,
\end{equation}
%
where the index $l$ is connected with a recursion level, i.e., with the coordinate number $i, 1 \leq i \leq N$, which this problem belongs to. In this case $x = y_i$, and the search domain $[a, b] = [a(l), b(l)]$ is the interval $[a_i, b_i]$. When $i = 1$, $f_l(x) = \varphi_1(y_1)$. If $i > 1$, then the index $l$ is also connected with the vector $\nu_{i - 1} = \nu_{i - 1}(l)$ and $f_l(x) = \varphi_i(\nu_{i - 1}, y_i)$. Moreover, any problem~\ref{eq:10} except the problem $f_l(x)$ is associated with the ``parent'' one, i.e., with a problem which has generated it.

Let us designate the operation of the computation of the objective function value at a given point of the search domain as the trial. Consider the methods, which generate during solving the problem~\ref{eq:10} the sequence of trial coordinates $x_l^1, x_l^2, \dots, x_l^k, \dots$ and the sequence of the trial results (the function values) $z_l^1, z_l^2, \dots, z_l^k, \dots$, where $z_l^j = f_l(x_l^j), j \geq 1$. After completing $k \geq 1$ trials the set of the search information is formed
\begin{equation}
\label{eq:11}
\omega_l^k = \big\{ (x_l^j, z_l^j, \lambda_l^j): 1 \leq j \leq k \big\},
\end{equation}
%
where $\lambda_l^j, 1 \leq j \leq k$, are the indices of the problems generated for the computation of the values $z_l^j, 1 \leq j \leq k$. Within the framework of the multidimensional scheme, it is reasonable to form the search information separately for each problem of the family $F_Q$ under solving. As a complete set of information for the multidimensional problem being solved one can consider the set
\begin{equation*}
\Omega_Q = \big\{ \omega_l^{k(l)}, 1 \leq l \leq Q \big\},
\end{equation*}
%
where $k(l), 1 \leq l \leq Q$, is the number of the search information elements for the problem with the index $l$. It is important to note that the values $z_l^j, 1 \leq j \leq k(l)$, in the search information for the function of the level less than $N$ may vary since these values are the estimates of the minimum values of the functions $f_l$, and these estimates may be refined during the computations.

Note that in traditional nested optimization scheme, in the course of solving a problem~\ref{eq:10} the information on the particular sets~\ref{eq:11} is used only while in the adaptive generalization the full set of information~\ref{eq:12} is taken into consideration.

Before the detailed description of the general adaptive scheme let us introduce some auxiliary terms. When a univariate subproblem generates for computing the value of its objective function a subproblem of the next level we will call the generating subproblem as parental one, or just a parent, and the generated subproblem as the upper one. Obviously, the subproblem~\ref{eq:9} of the first level can be parental only and subproblems of the level $N$ have no upper levels. Let us juxtapose to the subproblem with the index $l$ from the set $F_Q$ (except the subproblem~\ref{eq:9} with the index $l = 1$) the number $\pi(l)$ of its parent and the number $\rho(l)$ of a recursion level (number of coordinate) which this subproblem belongs to. Moreover, let us consider a point $x_l^j$ being $j$-th trial in the course of solving the problem~\ref{eq:10} with the number $l$ and connect with it the number $\lambda_l^j$ of a subproblem generated for computing the function value at this point. Let us introduce as well the designation $\varphi_l^*$ for the current minimal value of objective function $f_l$ from~\ref{eq:10}.

\begin{enumerate}[label=\bfseries Step \arabic*., leftmargin=*]
  \item \textbf{Initialization.}
  \begin{enumerate}[label=\arabic{enumi}.\arabic*., leftmargin=-2em]
    \item Set the number of subproblems of the nested scheme $Q = 0$ and the number of coordinate level $l = 1$. Accept the level number $i = 1$. Choose a point $x_1^1$ in the interval $[a_1, b_1]$ as the starting point of solving the univariate subproblem~\ref{eq:9}. Increase $Q$ by $1$ and include this problem in the set $F_Q$ with the index $l = 1$.

    \item For calculating at the level $i$ the value of the objective function at the point $x_i^1$, pass to the next coordinate ($i = i + 1$) for solving the univariate subproblem
    \begin{equation}
      \label{eq:12}
      \min_{y_i \in [a_i, b_i]} \varphi_i(\hat{\nu}_{i - 1}, y_i),
    \end{equation}
    where the components of the vector $\hat{\nu}_{i - 1} = (\nu_1, \dots, \nu_{i - 1})$ are constants. Namely, $\nu_j = x_j^1, 1 \leq j \leq i - 1$.

    \item Choose a point $x_i^1$ in the interval $[a_i, b_i]$ as the starting point for solving the subproblem~\ref{eq:12}. Increase $Q$ by $1$ and include this problem in the set $F_Q$ with the index $l = i$.

    \item Repeat operations 1.2 and 1.3 up to reaching the level $i = N$. At this level calculate the value $z_N^1$ of objective $\varphi(y)$ function from~\ref{eq:1} at the point $y = (x_1^1, \dots, x_N^1)$ instead of generating a new univariate subproblem.

    \item For all subproblems from $F_Q$ with numbers $l, 1 \leq l \leq Q$:
    \begin{enumerate}[label=\arabic{enumi}.\arabic{enumii}.\arabic*., leftmargin=*]
      \item If $l \leq N$, then set $\lambda_l^1 = l + 1$ (for the $l = N$ number $\lambda_l^1$ is undefined).
      \item Accept $z_l^1 = z_N^1$, $k(l) = 1$, and form the sets of search information $\omega_l^{k(l)} = \{( x_l^1, z_l^1, \lambda_l^1 )\}$ (except $l = N$ where $\omega_l^{k(l)} = \{( x_l^1, z_l^1 )\}$).
      \item Specify current minimal value $\varphi_l^* = z_l^1$, $\rho(l) = l$, and for $l > 1$ accept $\pi(l) = l - 1$.
    \end{enumerate}
  \end{enumerate}

  \vspace{3 mm}
  \item \textbf{Main procedure.}
  \begin{adjustwidth}{-4em}{}
    %Add margin%
    Let there be $Q$ subproblems~\ref{eq:10} for which sets~\ref{eq:11} have been formed and the values $\pi(l)$, $\rho(l)$, $\varphi_l^*$ have been determined. The next iteration of the adaptive scheme is executed in accordance with the following rules.
  \end{adjustwidth}
  \begin{enumerate}[label=\arabic{enumi}.\arabic*., leftmargin=-2em]
    \item Choice of univariate subproblem.
    \begin{enumerate}[label=\arabic{enumi}.\arabic{enumii}.\arabic*., leftmargin=*]
      \item Assign to each subproblem~\ref{eq:10} with the index $l, 1 \leq l \leq Q$, the value $\widetilde{R}(l)$ called the characteristic of this subproblem.
      \item Find among the subproblems~\ref{eq:10} a subproblem with index $\sigma, 1 \leq \sigma \leq Q$, which the maximal characteristic corresponds to
      \begin{equation}
      \label{eq:13}
      \widetilde{R}(\sigma) = \max \big\{ \widetilde{R}(l): 1 \leq l \leq Q \big\}.
      \end{equation}
    \end{enumerate}
    \item New iteration.
    \begin{enumerate}[label=\arabic{enumi}.\arabic{enumii}.\arabic*., leftmargin=*]
      \item Set $l = \sigma, i = \rho(\sigma)$ and determine a point $x_l^{k(l) + 1}$ of the next trial in the subproblem $\sigma$ taking into account the search information~\ref{eq:11}. If $i < N$, then accept $\lambda_l^{k(l) + 1} = Q + 1$.

      \item If $i = N$, then calculate the value $z_l^{k(l) + 1} = \varphi(\hat{\nu}_{i - 1}, x_l^{k(l) + 1})$ where the vector $\hat{\nu}_{i - 1}$ has been determined in preceding parental subproblems. Renew the current estimation of minimum $\varphi_l^*$ and the search information adding the pair $(x_l^{k(l) + 1}, z_l^{k(l) + 1})$ to the set~\ref{eq:11} and incrementing $k(l)$ by $1$. Then pass to Substep 2.3. In the opposite case ($i < N$) go to 2.2.3.

      \item For calculating at the level $i$ the value of the objective function at the point $x_l^{k(l) + 1}$ take the next coordinate ($i = i + 1$) for solving a new univariate subproblem~\ref{eq:12} (the components of the vector $\hat{\nu}_{i - 1}$ are fixed). Increase $Q$ by $1$ and assign to the new subproblem the number $l = Q$ in the set $F_Q$.

      \item Choose a point $x_l^1$ in the interval $[a_l, b_l]$ as the starting point for solving the new subproblem~\ref{eq:12}. Define the set~\ref{eq:11} as empty one and assign $k(l) = 0$. Accept $\varphi_l^* = +\infty$, $\rho(l) = i$, $\lambda_l^1 = Q + 1$.

      \item If the subproblem $\sigma$ is parental for the current one, then set $\pi(l) = \sigma$, otherwise, $\pi(l) = Q - 1$. Go to 2.2.2.
    \end{enumerate}

    \item Renewing the search information in parental subproblems.
    \begin{enumerate}[label=\arabic{enumi}.\arabic{enumii}.\arabic*., leftmargin=*]
      \item Return to the parental subproblem: new $l = \pi(l)$.

      \item If the set $\omega_l^{k(l)}$ is empty or $l = \sigma$, then accept the current minimal value of subproblem with number $\lambda_l^{k(l)}$ as the objective function value $z_l^{k(l)}$ at the new trial point $x_l^{k(l)}$, include in the set $\omega_l^{k(l)}$ the triplet $\big( x_l^{k(l)}, z_l^{k(l)}, \lambda_l^{k(l)} \big)$, add unity to $k(l)$ and renew the estimate $\varphi_l^*$.

      \item Repeat operations 2.3.1 and 2.3.2, if $l \neq \sigma$. In the opposite case, go to 2.3.4.

      \item Accept $\mu = \pi(l)$ and pass to the parental subproblem: new $l = \pi(l)$.

      \item Find in the set $\omega_l^{k(l)}$ a triplet $\big( x_l^j, z_l^j, \lambda_l^j \big), 1 \leq j \leq k(l)$, for which $\lambda_l^j = \mu$ and replace $z_l^j$ with $\varphi_{\mu}^*$.

      \item If $\varphi_{\mu}^* < \varphi_l^*$, then replace $\varphi_l^*$ with $\varphi_{\mu}^*$ and go to 2.3.4. In the opposite case, complete the current iteration of the main procedure and return to Substep 2.1.

    \end{enumerate}
  \end{enumerate}
\end{enumerate}

Description of the adaptive nested optimization scheme has been completed.

In the algorithmic scheme presented above the decision rules for choice of starting points in univariate subproblems and for selection the best subproblem (characteristics $\widetilde{R}(l)$ in 2.2.1) are not defined. Generally speaking, this characteristic $\widetilde{R}(l)$ is determined by a one-dimensional method that are used for solving the subproblem $l$.

In the present research, the global search algorithm developed within the framework of the information approach (see \cite{Ref34, Ref35, Ref36}) has been applied for solving the one-dimensional problems~\ref{eq:10}. Let us show how this method forms the subproblem characteristic $\widetilde{R}(l)$. According to this algorithm, while solving a univariate problem~\ref{eq:10} the first trial is executed at some internal point of the interval $[a, b]$. For certainty, the starting point $x_l^1$ is considered to be the middle of the interval $[a, b]$ where the value $z_l^1 = f_l(x_l^1)$ is computed.

The choice of the point $x_l^{k + 1}, k > 1$ of any subsequent $(k + 1)$-th trial is determined by the following operations:
\begin{enumerate}[label=\bfseries Step \arabic*., leftmargin=0em, itemindent=*, itemsep=1em]
  \item Coordinates of the previous trials $x_l^1, x_l^2, \dots, x_l^k$ from~\ref{eq:11} and the ends of the interval $[a, b]$ are renumbered by subscripts in increasing order:
  \begin{equation}
  \label{eq:14}
  a = x_0 < x_1 < \dots < x_k < x_{k + 1} = b,
  \end{equation}
  %
  and the values $z_j = f_l(x_j), 1 \leq j \leq k$, are juxtaposed to them (besides $x_0$ and $x_{k + 1}$). Values $z_j, 1 \leq j \leq k$, are taken from the set~\ref{eq:11} because they are the corresponding outcomes $z_j, 1 \leq j \leq k$.

  \item For each interval $(x_{j - 1}, x_j), 1 \leq j \leq k + 1, k = k(l)$, a value $R(j)$ hereinafter called the characteristic of the interval is calculated as follows

  \begin{equation*}
  R(1) = 2m(x_1 - x_0) - 4z_1, j = 1,
  \end{equation*}
  %
  \begin{equation}
  \label{eq:15}
  R(j) = m(x_j - x_{j - 1}) + \frac{(z_j - z_{j - 1})^2}{m(x_j - x_{j - 1})} - 2(z_j + z_{j - 1}), 1 < j < k + 1,
  \end{equation}
  %
  \begin{equation*}
  R(k + 1) = 2m(x_{k + 1} - x_k) - 4z_k, j = k + 1
  \end{equation*}
  %

  \item The interval $(x_{t - 1}, x_t)$ with the index $t = t(l)$, which the maximal characteristic $R(t)$ corresponds to is selected, i.e.,
  \begin{equation}
  \label{eq:16}
  R(t) = \max \big\{ R(j): 1 \leq j \leq k + 1 \big\},
  \end{equation}
  %
  and the value $R(t)$ is accepted as the characteristic $\widetilde{R}(l)$ of the subproblem $l$.

  \item The next trial points in the interval with the maximal characteristic is calculated as
  \begin{equation}
  \label{eq:17}
  x^{k + 1} = \frac{x_t + x_{t - 1}}{2} - \frac{z_t - z_{t - 1}}{2m}, 1 < t < k + 1,
  \end{equation}
  \begin{equation*}
  x^{k + 1} = \frac{x_t + x_{t - 1}}{2}, t = 1, t = k + 1.
  \end{equation*}

  Here the value $m$ can be considered as an estimate of Lipschitz constant of the minimized function. It can be calculated adaptively on the base of the trial results as
  \begin{equation}
  \label{eq:18}
    m =
      \begin{cases}
      rM, & \quad M > 0, \\
       1, & \quad M = 0
      \end{cases}
  \end{equation}
  %
  where
  \begin{equation*}
  M = \max_{1 < i \leq k} \frac{|z_i - z_{i - 1}|}{x_i - x_{i - 1}},
  \end{equation*}
  %
  and $r > 1$ is a parameter of the method.

  The algorithm can be supplemented with a termination criterion:

  \item If
  \begin{equation}
  \label{eq:19}
  x_t - x_{t - 1} \leq \varepsilon,
  \end{equation}
  where $t$ from~\ref{eq:16} is the index of the interval with the maximal characteristic, then stop. In the opposite case, it is necessary to calculate $z_l^{k + 1} = f_l(x_l^{k + 1})$, to increment $k$ by $1$ and to pass to Step 1. Here $\varepsilon > 0$ is the predefined search accuracy (for the coordinates).
\end{enumerate}
The univariate method describes above belongs to the class of the characteristical global search algorithm (see \cite{Ref11, Ref36}). This class includes many well known one-dimensional multiextremal optimization methods, such as, for example, algorithms~\cite{Ref18, Ref25, Ref26, Ref32, Ref36} which can be embedded into the adaptive nested scheme as well.

Using the rule~\ref{eq:19} one can introduce a termination condition into the adaptive nested scheme. It can be either a ``strong'' criterion when~\ref{eq:19} is satisfied for all problems from the set $F_Q$ or a weaker one when the inequality~\ref{eq:19} holds for the problem~\ref{eq:9} only.

\section{Convergence of the adaptive nested optimization scheme}
\label{sec:3}
In order to prove the convergence of the proposed approach, let us note first that in the novel generalized multistep scheme \textit{the current estimate of the minimum value} of the function $\widetilde{\varphi}_{i + 1}(y_{i + 1})$ determined on the base of the completed global search iterations is used as the sought function value $\widetilde{\varphi}_i(y_i)$. The estimates used can be refined during the computations. However, the estimates may differ from the values obtained by the exact solving of problems~\ref{eq:4} essentially. More strictly, one can say that in the global search the computations deal not with the ``exact'' one-dimensional functions $\widetilde{\varphi}_i(y_i), 1 \leq i < N$, but with the approximations $\widetilde{\psi}_i(y_i), 1 \leq i < N$, of these ones. Therefore, after completing the optimization, some estimate $\psi^*$ of the exact solution $\varphi^*$ from~\ref{eq:4} will be obtained. In this connection, an important question on the nearness of the estimates $\psi^*$ and $\varphi^*$ arises. It is worth noting that the approximate computations of the one-dimensional function values are inherent in the basic nested optimization scheme as well (see \cite{Ref34, Ref35}). For this scheme, the following theorem is true:

\newtheorem{mytheorem}{Theorem}
\begin{mytheorem}
\label{th:1}
% Remove italic
(see \cite{Ref34}) Let the functions $\varphi_i(\nu_i)$ be the Lipschitzian ones with the constants $L_i$ for any $i$, $1 \leq i \leq N$. Then, the following assertion is true for the approximation $\psi^*$ obtained in the minimization of the function $\varphi(y), y \in D$, by the information algorithm using the basic scheme of nested optimization:
\begin{equation*}
\lim_{\varepsilon_i \to 0} \psi^* = \varphi^*,
\end{equation*}
if
\begin{equation*}
\varepsilon_{i + 1} = O(\varepsilon_i), 1 \leq i < N,
\end{equation*}
and besides, the inequality
\begin{equation}
\label{eq:20}
m_i \geq 2L_i + \theta, 1 \leq i < N, \theta > 0,
\end{equation}
holds after terminating the minimization for coordinate $y_i$ (i.e., after solving any problem~\ref{eq:10} with $l = l(i)$).
\end{mytheorem}

Here $m_i$ are estimates~\ref{eq:18} applied in the solving of subproblems~\ref{eq:10} for the corresponding coordinates $y_i$, $\varepsilon_i$ are the accuracies for the solutions of these problems in termination criterion~\ref{eq:19}, and $O(\varepsilon_i)$ are the infinitesimal in comparison with $\varepsilon_i$.

Theorem~\origref{th:1} is a theoretical substantiation of the applicability of the basic nested optimization scheme. At the same time, it could be considered as the justification of the convergence of the adaptive scheme as well if a ``strong'' termination condition is applied, i.e., when~\ref{eq:19} is satisfied for all one-dimensional subproblems. In this case, the order of computations is changed in the realization of the scheme only.

In the case of using the adaptive scheme with the ``weak'' termination condition the applicability of the theorem is limited. The one-dimensional problems $f_l(x), 1 \leq l \leq Q,$ of the family $F_Q$ from~\ref{eq:10} are considered simultaneously. A part of these ones (with large enough values of the minimized function $\varphi(y)$) may be never solved with the required accuracy $\varepsilon$. This is the initial idea for improvement efficiency of the adaptive scheme. As a result, the error of the calculations of the functions $\varphi_i(\nu_i)$ may appear to be considerable. This error may remain unchanged even when $\varepsilon$ tends to zero. In order to resolve this problem, one can note that actually there is no need in the exact calculation of the functions $\varphi_i(\nu_i)$ in all points of the feasible domain. The necessary requirement is the condition of the exact calculation of these functions in the nearness of the global minimum of the initial function $\varphi(y)$ only. Just because this requirement will be used further for the theoretical substantiation of the adaptive nested optimization scheme.

Let us represent the search information $\omega_l^{k(l)}, 1 \leq l \leq Q$, from~\ref{eq:11}--\ref{eq:12} obtained during the global search for the functions $f_l(x) = \varphi_i(\nu_i, x), 1 \leq l \leq Q$, where $x = y_i, i = i(l)$, in the form
\begin{equation}
\label{eq:21}
\omega_l^{k(l)} = \{ (y_{l, j}, z_{l, j}(\delta_{l, j}), \lambda_{l, j}): 1 \leq j \leq k(l) \},
\end{equation}
%
where $y_{l, j}, 1 \leq j \leq k(l)$, are the points of the completed $l$ search trials, the index $l$ corresponds to the index of the problem in the family $F_Q$, the index $j$ corresponds to the increasing order of the iteration points~\ref{eq:14} in the search information, $\delta_{l, k}, 1 \leq j \leq k(l)$, are the errors of the computation of the values $z_{l, j}, 1 \leq j \leq k(l)$, i.e.,
\begin{equation}
\label{eq:22}
(z_{l, j} - \varphi_{\lambda_{l, j}}(\nu_{\lambda_{l, j}}, y_{l, j})) \leq \delta_{l, j}.
\end{equation}
%
Remind (see~\ref{eq:9}) that $\lambda_{l, j}, 1 \leq j \leq k(l)$, are the indices of the problems from the family $F_Q$, which were generated to calculate the values $z_{l, j}, 1 \leq j \leq k(l)$. It is important to emphasize also that the errors $\delta_{l, j}, 1 \leq j \leq k(l)$, could appear to be different for different points $z_{l, j}, 1 \leq j \leq k(l)$.

The possibility to obtain the estimates of the global minimum of the initial function $\varphi(y), y \in D$, is defined by the conditions of the following theorem.

\begin{mytheorem}
\label{th:2}
Let the adaptive scheme using the information global search algorithm~\ref{eq:14}--\ref{eq:18} and the ``weak'' termination condition~\ref{eq:19} for problem~\ref{eq:9} with $\varepsilon = 0$ be applied to solving problem~\ref{eq:1}. If
\begin{enumerate}[label=\roman*)]
  \item the objective function $\varphi(y)$ of problem~\ref{eq:1} satisfies the Lipschitz condition with a finite constant $L > 0$ in domain~\ref{eq:2};

  \item starting from some search step, the inequality
  \begin{equation}
  \label{eq:23}
  m > 2L,
  \end{equation}
  for the estimates $m$ of Lipschitz constant from~\ref{eq:18} holds;

  \item for any function $f_l(x) = \varphi_i(\nu_i, x), 1 \leq l \leq Q, i = i(l)$, of the family $F_Q$ from~\ref{eq:10} the following conditions for the errors of the function values' calculation hold
  \begin{multline}
  \label{eq:24}
  \delta_{l, \tau} + \delta_{l, \tau - 1} \leq 2\varphi(y^*) + L(y_{l, \tau} - y_{l, \tau - 1}) - \\
  - (\varphi_{\lambda_{l, \tau}}(\nu_i, y_{l, \tau}) + \varphi_{\lambda_{l, \tau - 1}}(\nu_i, y_{l, t - 1})), 1 < \tau \leq k(l),
  \end{multline}
  %
  \begin{equation}
  \label{eq:25}
  \delta_{l, \tau} \leq \varphi(y^*) + L(y_{l, \tau} - y_{l, \tau - 1}) - \varphi_{\lambda_{l, \tau}}(\nu_i, y_{l, \tau}), \tau = 1,
  \end{equation}
  %
  \begin{equation}
  \label{eq:26}
  \delta_{l, \tau - 1} \leq \varphi(y^*) + L(y_{l, \tau} - y_{l, \tau - 1}) - \varphi_{\lambda_{l, \tau - 1}}(\nu_i, y_{l, \tau - 1}), \tau = k + 1,
  \end{equation}
  %
  where $\tau, 1 \leq \tau \leq k(l) + 1$, is the index of the interval (determined in accordance with the search information $\omega_l^{k(l)}$) containing the value of the $i^{\text{th}}$ coordinate $y_i^*$ of the global minimum point, $y^*$ i.e.,
  \begin{equation}
  \label{eq:27}
  y_{l, \tau - 1} \leq y_i^* \leq y_{l, \tau}.
  \end{equation}
\end{enumerate}

Then, the global minimizer $y^*$ of the function $\varphi(y)$ in the domain $D$ is the limit (accumulation) point of the search trial sequence generated by the adaptive scheme.
\end{mytheorem}

\begin{proof}
Since in termination condition~\ref{eq:19} $\varepsilon = 0$, the algorithm of the adaptive scheme will generate an infinite sequence of trials $\{y^s\}$. Assume some global minimum point $y^*$ in problem~\ref{eq:1} not to be the limit one of this sequence, i.e., there is no subsequence of the sequence $\{y^s\}$ converging to $y^*$. Because $D$ is a bounded and closed set, a subsequence converging to a limit point $\bar{y} \in D$ can be always selected from the sequence $\{y^s\}$ and $\varphi(\bar{y}) \geq \varphi(y^*)$.

As it follows~\cite{Ref36} from the property of the bilateral convergence of the information algorithms for the characteristics $R(\cdot)$ from~\ref{eq:15}, for the intervals containing the values of the coordinates of the point $\bar{y}$ holds
\begin{equation}
\label{eq:28}
\lim_{q \to 0} R(t(q)) = -4\varphi(\bar{y}),
\end{equation}
%
where $t$ is the index of the interval containing the values of the coordinates of the point $\bar{y}$, and $q$ is the index of the global search iteration. Let us estimate the characteristic $R(\tau), \tau = \tau(q)$, of the interval~\ref{eq:27}.

If $1 \leq \tau \leq k(l)$, then according to~\ref{eq:22},~\ref{eq:24}
\begin{multline*}
z_{l, \tau - 1} + z_{l, \tau} \leq \varphi_{\lambda_{l, \tau - 1}}(\nu_i, y_{l, \tau - 1}) + \varphi_{\lambda_{l, \tau}}(\nu_i, y_{l, \tau}) + \delta_{l, \tau - 1} + \delta{l, \tau} \leq \\
\leq 2\varphi(y^*) + L(y_{l, \tau} - y_{l, \tau - 1})
\end{multline*}
%
However,
\begin{multline*}
R(\tau) = m(y_{l, \tau} - y_{l, \tau - 1}) + \frac{(z_{l, \tau} - z_{l, \tau - 1})^2}{m(y_{l, \tau} - y_{l, \tau - 1})} - 2(z_{l, \tau} + z_{l, \tau - 1}) \geq \\
\geq m(y_{l, \tau} - y_{l, \tau - 1}) - 2L(y_{l, \tau} - y_{l, \tau - 1}) - 4\varphi(y^*),
\end{multline*}
%
therefore because of~\ref{eq:23}
\begin{equation}
\label{eq:29}
R(\tau) > -4\varphi(y^*)
\end{equation}

If $\tau = 1$ then taking into account~\ref{eq:22},~\ref{eq:25}
\begin{equation*}
z_{l, \tau} \leq \varphi_{\lambda_{l, \tau}}(\nu_i, y_{l, \tau}) + \delta_{l, \tau} \leq \varphi(y^*) + L(y_{l, \tau} - y_{l, \tau - 1}),
\end{equation*}
%
the estimate
\begin{equation*}
R(\tau) \geq 2m(y_{l, \tau} - y_{l, \tau - 1}) - 4L(y_{l, \tau} - y_{l, \tau - 1}) - 4\varphi(y^*)
\end{equation*}
%
is true and because of~\ref{eq:23} inequality~\ref{eq:29} holds again. The analogous estimates leading to the truth of~\ref{eq:29} are valid in the case $\tau = k + 1$ as well.

Thus, taking into account that $\varphi(y^*) \leq \varphi(\bar{y})$ the expressions~\ref{eq:28} and~\ref{eq:29} for sufficiently great $s$ contradict the assumption that the point $\bar{y}$ is a limit one since the characteristics of the intervals containing this point starting from some iteration of the global search become less than the ones of the intervals with the global minimum point. According to rules~\ref{eq:13},~\ref{eq:16} such intervals cannot be selected to place the trials inside these ones.
\end{proof}

The proof has been completed.

Computational experiments on test classes of multidimensional multiextremal problems presented in the next section confirm the theoretical properties of the methods considered.

\section{Numerical experiments}
\label{sec:4}
First, let us compare the character of the trial placement by the basic variant of the nested optimization scheme and by the adaptive one for a function from the essentially multiextremal test class (see \cite{Ref10, Ref11}). The functions of this class have the form
\begin{multline}
\label{eq:30}
\varphi(y_1, y_2) = -\Bigg\{ \bigg( \sum_{i = 1}^{7}\sum_{j = 1}^{7} \Big[ A_{i j}a_{i j}(y_1, y_2) + B_{i j}b_{i j}(y_1, y_2) \Big] \bigg)^2 + \\
+ \bigg( \sum_{i = 1}^{7}\sum_{j = 1}^{7} \Big[ C_{i j}a_{i j}(y_1, y_2) + D_{i j}b_{i j}(y_1, y_2) \Big] \bigg)^2 \Bigg\}^\frac{1}{2}
\end{multline}
%
where $a_{i j}(y_1, y_2) = \sin(\pi i y_1)\sin(\pi j y_2)$ and $b_{i j}(y_1, y_2) = \cos(\pi i y_1)\cos(\pi j y_2)$ and are considered in the domain $y_1, y_2 \in [0, 1]$. The parameters
\begin{equation*}
A_{i j}, B_{i, j}, C_{i, j}, D_{i, j} \in [-1, 1],
\end{equation*}
are the independent random numbers, distributed uniformly over the interval specified above.

In Fig.~\origref{fig:2}, the level curves of a function from the class~\ref{eq:30} as well as the distributions of the trials (marked by dark dots) in the course of optimization executed by the basic and adaptive schemes are plotted. The numerical experiment has been carried out for both variants with the parameter of the method $r = 3$ in the estimate~\ref{eq:18} and the search accuracy in termination condition~\ref{eq:19} $\varepsilon = 0.01$. Note that according to the basic scheme, condition~\ref{eq:18} should be satisfied for each one-dimensional subproblem while in the adaptive generalization for subproblem~\ref{eq:9} only.

Both adaptive and basic schemes have provided the required accuracy of the problem solution. However, the adaptive variant has carried out $257$ search trials while the basic scheme has computed $1139$ objective function values.

\begin{figure}[H]
\includegraphics[width=116mm]{Fig2.png}
\caption{\label{fig:2}The placement of the trials by the adaptive scheme (the left panel) and by the basic one (the right panel).}
\end{figure}

The comparison of the efficiencies of the optimization algorithms using the examples of solving particular test problems, obviously, is not a sufficiently convincing way for proving the advantages of one or another method. Carrying out some multiple experiments in order to obtain the statistically reliable results of the comparison is a more objective approach. For this goal one can use the method of operating  characteristics introduced in \cite{Ref10, Ref36}.

The \textit{operating characteristic} is a set of pairs:
\begin{equation*}
\{ \langle k, p(k) \rangle \},
\end{equation*}
%
where $k$ is the number of search iterations and $p(k)$ is the number of the problems from the test class solved successfully in $k$ iterations. An individual pair corresponds to a particular set of the parameters of the method, which the test problems were solved by. Such indicators can be calculated according to the results of numerical experiments for various parameters of the methods and presented as a plot in the $(k, p)$ axes. If for a given value $k$ the operating characteristic of a method is situated  above the  characteristic of another method, the first method provides a better reliability, i.e., the possibility of the proper solution of the problem at the same computational costs for the search. If for a given value $p(k)$ the operating characteristic of a method is located on the left of the characteristic of another method, the first method requires less computational expenditures to achieve the same reliability. Thus, the operating characteristics allow comparing the efficiencies of various methods visually.

Let us first consider the class of two-dimensional functions~\ref{eq:30} consisting of $100$ functions. For a comparative experiment four global optimization methods have been taken. These are the basic nested optimization scheme (BS), the adaptive one (AS) and two methods based on other approaches to solving the global optimization problems: the partitioning method DIRECT (see~\cite{Ref17}) and the algorithm applying the dimensionality reduction by means of Peano mappings (PMA) from~\cite{Ref36}. In BS and AS the reliability parameter value $r = 2.6$ was used, for PMA $r = 3.2$ (these values provide sufficient conditions of convergence to global minimum for these methods). The operating characteristics have been obtained by varying the accuracy in the termination condition~\ref{eq:19} for AS, BS and PMA and the number of trials as the termination rule in DIRECT. The problem was considered to have been solved successfully if the coordinates of the minimal value found in the course of optimization are located in the circular neighborhood of the real minimum with the radius $0.01$. The results of the experiment are presented in Table~\origref{tab:1} and Fig.~\origref{fig:3}. The number of trials is plotted on the abscissa axis in the logarithmic scale.

\begin{table}[H]
\centering
\caption{\label{tab:1}The number of the problems solved successfully in dependence on trials spent.}
\begin{tabular}{|c|c|c|c|c|}
  \hline
  Number of trials &  AS &  BS & PMA & DIRECT \\ \hline
  60               &  40 &  42 &  25 & 13     \\ \hline
  100              &  51 &  61 &  52 & 31     \\ \hline
  150              &  72 &  73 &  71 & 52     \\ \hline
  200              &  87 &  84 &  80 & 66     \\ \hline
  250              &  88 &  87 &  88 & 75     \\ \hline
  300              &  89 &  90 &  92 & 83     \\ \hline
  400              &  94 &  94 &  96 & 89     \\ \hline
  500              &  99 &  96 &  96 & 92     \\ \hline
  700              & 100 &  98 &  97 & 94     \\ \hline
  800              & 100 &  99 &  99 & 97     \\ \hline
  1000             & 100 &  99 & 100 & 99     \\ \hline
  1100             & 100 & 100 & 100 & 100    \\ \hline
\end{tabular}
\end{table}

\begin{figure}[H]
\includegraphics[width=120mm]{Fig3.png}
\caption{\label{fig:3}Operating characteristics on the class~\ref{eq:30}.}
\end{figure}

The numerical experiments have demonstrated an appreciable advantage of the adaptive scheme over the basic prototype and better performance of the adaptive scheme in providing a better reliability in comparison with the method DIRECT and the algorithm based on Peano curves.

The second series of the numerical experiments has been carried out using a well-known class of test functions GKLS from \cite{Ref6} for the 5-dimensional problems. The testing has been realized for a sample set of $100$ GKLS functions belonging to the most complicated GKLS class among the test classes described in~\cite{Ref6}.

Fig. \origref{fig:4} presents the operating characteristics for the basic nested optimization scheme (BS), the adaptive scheme (AS) and the Peano-mapping-based algorithm (PMA). All the methods used the value of parameter $r = 5$. The choice of the value $r = 5$ was caused by the necessity to satisfy the sufficient conditions of convergence for PMA. For this method, the inequality $m > 4L$ is the sufficient condition for the global convergence from \cite{Ref32}. The number of trials is plotted on the abscissa axis in the logarithmic scale.
\begin{figure}[H]
\includegraphics[width=116mm]{Fig4.png}
\caption{\label{fig:4}Operating characteristics using GKLS class.}
\end{figure}

As it follows from Fig. \origref{fig:4}, performance of the method BS is significantly worth than performance of the algorithms AS and PMA. The explanation of this fact consists in the phenomenon of the information loss by BS in the course of optimization as it was discussed above. Concerning the comparison of AS and PMA, one can note that for small number of iterations (and, as a result, for low reliability) the method using Peano  curves  has some  advantage. However, for achievement of a high reliability AS is more efficient than BS because it spends less number of trials.
\section{Conclusions}
\label{sec:7}
The work has been done in the framework of global optimization approach, according to which in order to construct the efficient global search algorithms, various dimensionality reduction schemes are applied. The reduction of the multidimensional problems to the one-dimensional ones allows using a wide choice of the one-dimensional algorithms to solve the multidimensional optimization problems.

In the present article, the nested optimization scheme of dimensionality reduction is considered. Along with a relative simplicity in its basic form, this scheme possesses a number of disadvantages including the redundancy of the number of computations of the minimized function values in the subdomains of the search domain. The authors have proposed a novel method to increase the efficiency of the initial scheme by means of using the full information on the problem in the course of global search. The novel adaptive scheme has been described in the generalized characteristic form, within the framework of which many one-dimensional global optimization algorithms can be employed. The generalized form has been specified for the case of the information global search method, which the proving of the convergence conditions has been done for. In order to confirm the efficiency of the developed approach, the results of the numerical experiments are presented justifying an essential acceleration of the search for the adaptive generalization as compared to the basic prototype and demonstrating the efficiency of the adaptive scheme as compared to the global optimization method based on application of Peano space-filling curves and to the popular method DIRECT.

It is worth noting that because of using the characteristic decision rule for selection of the coordinates for the next iteration, the adaptive scheme has a significant potential for the parallelization that may be a promising direction for its further development.

\begin{acknowledgements}
The authors thank Ruslan Israfilov for a significant contribution to computational experiments. The revised version of the paper was supported by the Russian Science Foundation, project  No. 15-11-30022 ``Global optimization, supercomputing computations, and applications''.
\end{acknowledgements}

\begin{thebibliography}{}
\bibitem{Ref1}
Butz, A.R.: Space-Filling Curves and Mathematical Programming. Inform. Control 12(4), 314--330 (1968)

\bibitem{Ref2}
Carr, C.R., Howe, C.W.: Quantitative Decision Procedures in Management and Economic: Deterministic Theory and Applications. McGraw–Hill, New York (1964)

\bibitem{Ref3}
Dam, E.R., Husslage, B., Hertog, D.: One-dimensional Nested Maximin Designs. Journal of Global Optimization 46(2), 287--306 (2010)

\bibitem{Ref4}
Floudas, C.A., Gounaris, C.E.: A Review of Recent Advances in Global Optimization. Journal of Global Optimization 45(1), 3--38 (2009)

\bibitem{Ref5}
Floudas, C.A., Pardalos, P.M., editors. Encyclopedia of Optimization. Kluwer Academic Publishers, Dordrecht (2001) (The 2nd ed. Springer, 2009)

\bibitem{Ref6}
Gaviano, M., Kvasov D.E., Lera D., Sergeyev Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM TOMS 29 (4), 469--480 (2003)

\bibitem{Ref7}
Gergel, V.P., Strongin, R.G.: Parallel computing for globally optimal decision making. Lecture Notes in Computer Science (including subseries Lecture Notes in Artificial Intelligence and Lecture Notes in Bioinformatics) 2763, 76-88 (2003)

\bibitem{Ref8}
Gergel, V., Grishagin, V., Israfilov R.: Local Tuning in Nested Scheme of Global Optimization. Procedia Computer Science 51, 865–874 (2015)

\bibitem{Ref9}
Goertzel, B.: Global Optimization with Space-Filling Curves. Applied Mathematics Letters 12(8), 133--135 (1999)`

\bibitem{Ref10}
Grishagin, V.A.: Operating Characteristics of Some Global Search Algorithms. Problems of Statistical Optimization 7, 198--206. Zinatne, Riga (1978) (In Russian)

\bibitem{Ref11}
Grishagin, V.A., Sergeyev, Y.D., Strongin, R.G.: Parallel Characteristic Algorithms for Solving Problems of Global Optimization. Journal of Global Optimization 10(2), 185--206 (1997)

\bibitem{Ref12}
Grishagin, V.A., Strongin, R.G.: Optimization of multiextremal functions subject to monotonically unimodal constraints. Engineering Cybernetics 5, 117--122 (1984)

\bibitem{Ref13}
Hime, A. E., Oliveira Jr., H.A., Petraglia, A.: Global Optimization Using Space-Filling Curves and Measure-Preserving Transformations. Soft Computing in Industrial Applications 96, 121--130 (2011)

\bibitem{Ref14}
Horst, R., Pardalos, P.M., editors.: Handbook of Global Optimization. Kluwer Academic Publishers, Dordrecht (1995)

\bibitem{Ref15}
Horst, R., Pardalos, P.M., Thoai, N.V.: Introduction to Global Optimization. Kluwer Academic Publishers, Dordrecht (1995). (2nd ed. Kluwer Academic Publishers, 2001).

\bibitem{Ref16}
Horst, R., Tuy, H.: Global Optimization: Deterministic Approaches. Springer-Verlag, Berlin (1990)

\bibitem{Ref17}
Jones, D.R.: The DIRECT Global Optimization Algorithm. In: Floudas C.A., Pardalos P.M. (eds) Encyclopedia of optimization 1. Kluwer Academic Publishers, Dordrecht, 431--440 (2001)

\bibitem{Ref18}
Kushner, H.J.: A new method of locating the maximum point of an arbitrary multipeak curve in the presence of noise. Transactions of ASME, Ser. D. Journal of Basic Engineering 86, 1, 97--106 (1964)

\bibitem{Ref19}
Kvasov, D.E., Pizzuti, C., Sergeyev, Y.D.: Local tuning and partition strategies for diagonal GO methods. Numerische Matematik 94(1),  93--106  (2003)

\bibitem{Ref20}
Kvasov, D.E., Sergeyev, Ya.D.: Multidimensional Global Optimization Algorithm Based on Adaptive Diagonal Curves. Comput. Math. Math. Phys. 43(1), 40--56 (2003)

\bibitem{Ref21}
Kvasov, D.E., Sergeyev, Y.D.: Lipschitz gradients for global optimization in a one-point-based partitioning scheme.  Journal of Computational and Applied Mathematics  236(16),  4042--4054 (2012)

\bibitem{Ref22}
Kvasov, D.E., Sergeyev, Ya.D.: Deterministic approaches for solving practical black-box global optimization problems. Advances in Engineering Software 80, 58-66 (2015)

\bibitem{Ref23}
Paulavicius, R., Sergeyev, Y.D., Kvasov, D.E., Zilinskas, J.: Globally-biased DISIMPL algorithm for expensive global optimization. Journal of Global Optimization 59(2--3), 545--567 (2014)

\bibitem{Ref24}
Pintér, J. D.: Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications). Kluwer Academic Publishers, Dordrecht (1996)

\bibitem{Ref25}
Piyavskij, S.A.: An Algorithm for Finding the Absolute Extremum of a Function. USSR Comput. Math. Math. Phys. 12(4), 57--67 (1972)

\bibitem{Ref26}
Sergeyev, Y.D., Grishagin V.A.: A Parallel Method for Finding the Global Minimum of Univariate Functions. Journal of Optimization Theory and Applications  80(3), 513--536 (1994)

\bibitem{Ref27}
Sergeyev, Y.D.: An efficient strategy for adaptive partition of N-dimensional intervals in the framework of diagonal algorithms. Journal of Optimization Theory and Applications 107(1), 145-168 (2000).

\bibitem{Ref28}
Sergeyev, Y.D., Grishagin, V.A.: Parallel Asynchronous Global Search and the Nested Optimization Scheme. J. Comput. Anal. Appl.  3(2), 123--145 (2001)

\bibitem{Ref29}
Sergeyev Ya.D.: Efficient partition of N-dimensional intervals in the framework of one-point-based algorithms. Journal of Optimization Theory and Applications, 124(2), 503-510 (2005).

\bibitem{Ref30}
Sergeyev, Ya. D., Strongin, R. G., Lera, D.: Introduction to Global Optimization Exploiting Space-Filling Curves. Springer (2013)

\bibitem{Ref31}
Shi, L., Ólafsson, S.: Nested Partitions Method for Global Optimization. Operations Research 48(3), 390--407 (2000)

\bibitem{Ref32}
Shubert, B.O.: A Sequential Method Seeking the Global Maximum of a Function. SIAM Journal on Numerical Analysis 9(3), 379--388 (1972)

\bibitem{Ref33}
Strekalovsky, A.S.: On Solving Optimization Problems with Hidden Nonconvex Structures. In: Rassias Th.M., Floudas Ch.A., Butenko S. (eds) Optimization and Science in Engineering. Springer, 465-502 (2014)

\bibitem{Ref34}
Strongin, R.G.: Numerical Methods in Multiextremal Problems (Information-Statistical Algorithms). Nauka, Moscow (1978). (In Russian)

\bibitem{Ref35}
Strongin, R.G.: The information approach to multiextremal optimization problems. Stochastics and Stochastic Reports 27, 65--82 (1989)

\bibitem{Ref36}
Strongin, R.G., Sergeyev, Ya.D.: Global Optimization with Non-convex Constraints: Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000, 2nd ed. 2013, 3rd ed. 2014)

\bibitem{Ref37}
Törn, A., Žilinskas, A.: Global Optimization. Lecture Notes in Computer Science 350. Springer-Verlag, Berlin (1989)

\bibitem{Ref38}
Zhigljavsky, A.A.: Theory of Global Random Search. Kluwer Academic Publishers, Dordrecht (1991)

\bibitem{Ref39}
Zhigljavsky, A.A., Chekmasov, M.V.: Comparison of independent, stratified and random covering sample schemes in optimization problems. Mathematical  and Computer Modelling  23(8--9), 97--110 (1996)
\end{thebibliography}

\end{document}

### 2020-06-13 2020/OPTIMA/optima_paper.tex
OPTIMA: Use a bibliography file

@@
 \usepackage{subfig}
+\usepackage{breqn}
 \usepackage[labelfont=bf,labelsep=period]{caption}
@@
 \inputencoding{utf8}
-\usepackage[english, russian]{babel}
+\usepackage[english]{babel}
 
@@
 %
-\title{Parallel global optimization algorithm for obtaining uniform convergence when simultaneously solving a set of global optimization problems\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}}
+\title{Parallel global optimization algorithm for obtaining uniform convergence when
+simultaneously solving a set of global optimization problems\thanks{This study was supported
+by the Russian Science Foundation, project No.\,16-11-10150.}}
 %
@@
 \begin{abstract}
-    В данной работе рассматривается построение параллельной версии алгоритма глобальной оптимизации, решающего одновременно множество задач с нелинейными ограничениями и получающего при этом равномерные оценки решений на этом множестве.
-Последнее свойство позволяет наиболее оптимально распределять вычислительные ресурсы, т.
-к.
-в процессе работы алгоритма погрешности численного решения во всех задачах убывают примерно с одинаковой скоростью.
-Алгоритм присваивает приоритет каждой задаче и на каждой итерации производит вычисления целевых функций в нескольких задачах параллельно.
-При окончании работы метода в произвольный момент времени во всех задачах из решаемой серии будут получены решения сходного качества.
-Серии из нескольких задач возникают, если задача глобальной оптимизации имеет дискретный параметр или, например, при решении задачи многокритериальной оптимизации методом свертки критериев.
-Рассматриваемый алгоритм использует отображения типа кривой Пеано для редукции многомерных задач оптимизации к одномерным.
-Эффективность реализованного алгоритма протестирована на наборах искусственно сгенерированных задач глобальной оптимизации, а также при решении серии задач, полученной в результате скаляризации задачи многокритериальной оптимизации.
-Также экспериментально подтверждена равномерная сходимость метода.
+In the present work, an extension of global optimization algorithm solving a number of problems
+simultaneously onto a case of problems with non-linear constraints is considered. The property
+of uniform convergence inherent to the algorithm allows distributing the computation resources
+in the optimal way since the errors of numerical solutions decrease with approximately equal
+rates in all problems in the course of the execution of the algorithm. The algorithm assigns a
+priority to each problem and performs the computations of the objective functions in several
+problems in parallel at every iteration. Upon stop of the method at arbitrary time moment,
+solutions of similar quality would be obtained in all problems of the series being solved. A series
+of several problems arises if the global optimization problem has a discrete parameter or, for
+example, in solving a multicriteria optimization problem using criteria convolution. The
+considered algorithm employs mappings like Peano space filling curve for the reduction of the
+multidimensional optimization problems to the one-dimensional ones.  The efficiency of the
+implemented algorithm was tested using the sets of artificially generated global optimization
+problems as well as in solving a series of problems obtained as a result of scalarization of a
+multicriteria optimization  problem. Also, the numerical experiments confirmed the uniform
+convergence of the method.
 
-\keywords{Global optimization \and Uniform convergence \and Derivative-free optimization \and Parallel computations.}
+\keywords{Global optimization \and Uniform convergence \and Derivative-free optimization
+\and Parallel computations.}
 \end{abstract}
@@
 
-\section{Введение}
+\section{Introduction}
 \label{sec-intro}
-Нелинейная глобальная оптимизация невыпуклых функций традиционно считается одной из самых трудных
-задач математического программирования. Отыскание глобального минимума функции от нескольких переменных
-зачастую оказывается сложнее, чем локальная оптимизация в тысячемерном пространстве. Для последней может оказаться достаточно
-применения простейшего метода градиентного спуска, в то время как чтобы \textit{гарантированно} отыскать глобальный оптимум методам
-оптимизации приходится накапливать информацию о поведении целевой функции во всей области поиска \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}.
-Решение серии таких задач при ограниченных вычислительных
-ресурсах является еще более сложной задачей: помимо поиска глобального экстремума необходимо
-распределять вычислительные ресурсы так, чтобы сразу во всех решаемых задачах положение глобального
-экстремума было оценено примерно с одинаковым качеством. Обычно серию из \(q\) задач решают либо последовательно, либо
-параллельно порциями по \(p\ll q\) задач, где \(p\) --- количество параллельных вычислительных устройств.
-Такой подход ведет к тому, что в каждый момент времени до окончания вычислений
-остаются задачи, в которых оценка глобального оптимума не получена вообще, в то время, как в задачах из начала
-списка оптимум может быть оценен даже с избыточной точностью.
+Nonlinear global optimization of the non-convex functions is considered traditionally to be one
+of the most difficult problems of mathematical programming.
+ The finding of the global minimum of a function of several variables appears to be more
+difficult often than a local optimization in a thousand-dimensional space.
+For the latter, application of the simplest gradient descent method may appear to be enough
+whereas in order to \textit{guarantee} the finding of the global optimum, the optimization
+methods have to accumulate the information on the behavior of the objective function in the
+whole search domain \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}.
+The solving of a series of such problems at limited computational resources is even more
+difficult problem: besides the search of the global extremum, it is necessary to distribute the
+computational resources in such a way that the location of the global extremum would be
+estimated with approximately equal quality in all problems being solved at once.
+Usually, a series of \(q\) problems is solved either sequentially or in parallel by portions of \(p\ll
+q\) problems where \(p\) is a number of parallel computational devices.
+Such an approach leads to a situation when at every time moment until the stop of computations
+some problems remain, in which the estimate of the global optimum is not obtained yet whereas
+in the problems from the beginning of the list the optimum may be estimated even with excess
+precision.
 
-В данной работе рассматривается обобщение ранее разработанного в ННГУ им. Н. И. Лобачевского
-параллельного метода глобальной оптимизации для одновременного решения множества задач \cite{BarkalovStrongin2018} на
-случай задач с нелинейными ограничениями. Для учета ограничений используется индексная схема \cite{Strongin2000},
-позволяющая работать с частично вычислимыми целевым функциями и обладающая экономичностью,
-сравнимой с другими подходами \cite{BarkalovLebedev2017}. Эффективность реализованного
-алгоритма показана на примере решения множеств задач, сгенерированных специализированным
-механизмом, порождающим наборы задач заданной размерности с заданным количеством нелинейных ограничений \cite{GergelBarkalov2019}.
-Кроме искусственно сгенерированных задач, рассматриваемый метод протестирован также
-на множестве задач, возникающем при решении задачи многокритериальной оптимизации
-с нелинейными ограничениями методом свертки критериев \cite{Ehrgott2005}.
+In the present work, a generalization of parallel global optimization method for simultaneous
+solving of a set of problems developed earlier in Lobachevskii University of Nizhnii Novgorod
+\cite{BarkalovStrongin2018} onto the case of nonlinear constraints is considered. To account
+for the constraints, the index scheme \cite{Strongin2000} is applied allowing working with
+partially computable objective functions and featured by efficiency comparable to other
+approaches \cite{BarkalovLebedev2017}. The convergence of the algorithm to the global
+optimal solutions in all problems was proved theoretically, the efficiency of the implemented
+algorithm was demonstrated by solving sets of problems generated by a special mechanism
+generating the sets of problems of predefined dimensionality with predefined number of
+nonlinear constraints \cite{GergelBarkalov2019}.
+Besides the problems generated artificially, the considered method was tested also on a set of
+problems arising when solving a multicriteria optimization problem with nonlinear constraints by
+criteria convolution method \cite{Ehrgott2005}.
 
-Статья имеет следующую структуру. В разделе \ref{sec:problem} рассмотрена
-постановка решаемой задачи, далее в разделе \ref{sec:method} приведено описание
-параллельного метода оптимизациии. Результаты численных экспериментов, подтверждающие
-эффективность рассматриваемого метода приводятся в разделе \ref{sec:exps}.
-В заключении приводится краткая сводка результатов, полученных в ходе работы, и
-указаны направления дальнейших усилий по улучшению программной реализации рассмотренного метода.
+The paper is structured as follows. In Section \ref{sec:problem}, the statement of the problem to
+be solved is considered. Then, in Section \ref{sec:method}, a description of the parallel
+optimization method is given. The sufficient conditions of convergence of the considered
+method are given in Section \ref{sec:conv_method}.
+The results of the numerical experiments confirming the efficiency of considered method are
+presented in Section \ref{sec:exps}.
+In Conclusion, a short summary of the results obtained in the present study is given and the
+directions of further effort on the improvement of the software implementation of the
+considered method are outlined.
 
-\section{Постановка задачи глобальной оптимизации}
+\section{Global optimization problem statement}
 \label{sec:problem}
-В рамках данной работы будем рассматривать следующую постановку задачи глобальной
-оптимизации: найти глобальный минимум \(N\)-мерной функции \(\varphi(y)\) в гиперинтервале
+Within the frames of the present work, we will consider the following global optimization
+problem statement: to find the global minimum of an \(N\)-dimensional function \(\varphi(y)\) in
+a hyperinterval
 \(D=\{y\in \mathbf{R}^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
-Для построения оценки глобального минимума по конечному количеству вычислений
-значения функции требуется, чтобы скорость изменения \(\varphi(y)\) в \(D\) была ограничена.
-В качестве такого ограничения как правило принимается условие Липшица.
+In order to construct an estimate of the global minimum based on a finite number of
+computations of the function values, the rate of variation of \(\varphi(y)\) in \(D\) is required to
+be limited.
+As a rule, the Lipschitz condition Is accepted as such a limitation.
 \begin{equation}
@@
 
-Существуют различные методы, решающие рассмотренную многомерную задачу напрямую \cite{SergeyevKvasov2017, Jones2009},
-а также эффективные методы решения одномерных задач \cite{Norkin1992, Strongin2000}. В данной работе рассматривается одномерный метод,
-который применяется совместно со схемой редукции размерности.
-Классической схемой редукции размерности исходной задачи для алгоритмов глобальной оптимизации является
-использование разверток --- кривых, заполняющих пространство \cite{Sergeyev2013}.
+There are various methods solving the considered multidimensional problem directly
+\cite{SergeyevKvasov2017, Jones2009} as well as efficient methods of solving the one-
+dimensional problems \cite{Norkin1992, Strongin2000}.
+In the present work, a one-dimensional method is considered, which is applied together with the
+dimensionality reduction scheme.
+The use of evolvents or space-filling curves is a classical scheme of reduction of the
+dimensionality of the initial problem for the global optimization algorithms
+\cite{Sergeyev2013}.
 \begin{equation}
 \label{cube}
-\lbrace y\in \mathbf{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
+\lbrace y\in \mathbf{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant
+N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
 \end{equation}
 
-Отображение вида (\ref{cube}) позволяет свести задачу в многомерном пространстве к решению
-одномерной ценой ухудшения ее свойств. В частности, одномерная функция \(\varphi(y(x))\)
-является не Липшицевой, а Гёльдеровой:
+A mapping of the kind (\ref{cube}) allows reducing a problem in the multidimensional space to
+solving a one-dimensional one at the expense of worsening its properties. In particular, the one-
+dimensional function \(\varphi(y(x))\) is not a Lipschitz but a H\"{o}lder one:
 \begin{displaymath}
@@
 \end{displaymath}
-где константа Гельдера \(H\) связана с константой Липшица \(L\) соотношением
+where the H\"{o}lder constant \(H\) is related to the Lipschitz one \(L\) by the relation
 \begin{displaymath}
-H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}.
+  H=2L\sqrt{N+3}.
 \end{displaymath}
 
-Область \(D\) также может быть задана с помощью функциональных ограничений, что
-значительно усложняет задачу.
-Постановка задачи глобальной оптимизации в этом случае будет иметь следующий вид:
+The domain \(D\) can also be defined by the constraint functions that complicates the problem
+essentially.
+The statement of the global optimization problem in this case will take the following form:
 \begin{equation}
   \label{eq:constrained_problem}
-  \varphi(y^*)=\min\{\varphi(y):g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}
+  \varphi(y^*)=\min\{\varphi(y):y\in G=D \cap \{y: g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}\}
 \end{equation}
-Обозначим \(g_{m+1}(y)=\varphi(y)\). Далее будем предполагать, что все функции \(g_k(y),1\leqslant k \leqslant m+1\)
-удовлетворяют условию Липшица в некотором гиперинтервале, включающем \(D\).
 
-Далее будем интересоваться решением серии из \(q\) задач вида (\ref{eq:constrained_problem}):
+Let us denote \(g_{m+1}(y)=\varphi(y)\). Hereafter, we will assume all functions
+\(g_k(y),1\leqslant k \leqslant m+1\)
+to satisfy the Lipschitz condition in some hyperinterval including \(D\).
+
+Further, let us consider solving a series of \(q\) problems of the kind
+(\ref{eq:constrained_problem}):
 \begin{equation}
   \label{eq:many_problems}
-  \min\left\{\varphi_1(y), y\in D_1 \right\}, \min\left\{\varphi_2(y), y\in D_2\right\},..., \min\left\{\varphi_q(y), y\in D_q\right\}.
+  \min\left\{\varphi_1(y), y\in G_1 \right\}, \min\left\{\varphi_2(y), y\in G_2\right\},...,
+\min\left\{\varphi_q(y), y\in G_q\right\}.
 \end{equation}
 
-\section{Описание метода глобальной оптимизации}
+\section{Description of the global optimization method}
 \label{sec:method}
 
-Принимая во внимание схему редукции размерности (\ref{cube}), будем при описании метода считать, что
-требуется найти глобальный минимум функции \(\varphi(x), x\in[0;1]\),
-удовлетворяющей условию Гёльдера, при ограничениях \(g_j(x)\), также
-удовлетворяющих этому условию на интервале \([0;1]\).
+Taking into account the dimensionality reduction scheme (\ref{cube}), let us assume when
+describing the method that it is required to find the global minimum of the function \(\varphi(x),
+x\in[0;1]\) satisfying the H\"{o}lder condition with the constraints \(g_j(x)\) also satisfying this
+condition in the interval \([0;1]\).
 
-Рассматриваемый индексный алгоритм глобального поиска (ИАГП) для решения
-одномерной задачи (\ref{eq:constrained_problem}) предполагает построение последовательности
-точек \(x_k\), в которых вычисляются значения минимизируемой функции или ограничений \(z_k = g_s(x_k)\).
-Для учета последних используется индексная схема \cite{Strongin2000}. Пусть \(Q_0=[0;1]\). Ограничение, имеющее номер
- \(j\), выполняется во всех точках области
+The index algorithm of global search (IAGS) for solving the one-dimensional problems
+(\ref{eq:constrained_problem}) considered here implies the constructing of a sequence of points
+\(x_k\), which the values of the minimized function or of the constraints \(z_k = g_s(x_k)\) are
+computed in.
+To account for the latter, the index scheme \cite{Strongin2000} was used.
+Let \(Q_0=[0;1]\). The constraint having the index \(j\) is satisfied in all points of the domain
 \begin{displaymath}
@@
 \end{displaymath}
-которая называется допустимой для этого ограничения. При этом допустимая область \(D\)
-исходной задачи определяется равенством: \(D=\cap _{j=0}^{m}Q_{j}\).
-Испытание в точке \(x\in [0;1]\) состоит в последовательном вычислении значений
-величин \(g_{1}(x),...,g_{\nu }(x)\), где значение индекса \(\nu\) определяется условиями:
-\(x\in Q_{j},0\leqslant j<\nu ,x\notin Q_{\nu }\). Выявление первого нарушенного ограничения
-прерывает испытание в точке \(x\). В случае, когда точка \(x\)  допустима, т. е.
-\(x\in D\) испытание включает в себя вычисление всех функций задачи. При этом значение
-индекса принимается равным величине \(\nu =m+1\). Пара \(\nu =\nu (x),z=g_{\nu }(x)\),
-где индекс \(\nu\) лежит в границах \(1\leqslant \nu \leqslant m+1\), называется результатом
-испытания в точке \(x\).
+which is called feasible for this constraint.
+At that, the feasible domain \(D\) of the initial problem is defined by the equality:
+ \(D=\cap _{j=0}^{m}Q_{j}\).
+A trial in the point \(x\in [0;1]\) consists in a sequence computing of the values
+\(g_{1}(x),...,g_{\nu }(x)\)
+where the values of the indices \(\nu\) are defined by the conditions:
+\(x\in Q_{j},0\leqslant j<\nu ,x\notin Q_{\nu }\).
+The finding of the first violation of any constraint terminates the trial in the point \(x\).
+In the case when the point \(x\) is feasible i. e. \(x\in D\), the trial includes the computing of all
+functions of the problem.
+At that, the value of the index is accepted to be equal to the magnitude \(\nu =m+1\).
+The pair \(\nu =\nu (x), z=g_{\nu }(x)\) where the index \(\nu\) falls into the borders \(1\leqslant
+\nu \leqslant m+1\) is called the result of the trial in the point \(x\).
 
-Такой подход к проведению испытаний позволяет свести исходную задачу с функциональными
-ограничениями к безусловной задаче минимизации разрывной функции:
+Such an approach to conducting the trials allows reducing the initial problem with functional
+constraints to an unconstrained problem of minimization of a discontinuous function:
 
@@
     \psi (x^{*})=\min_{x\in [0;1]}\psi (x), \\
-    \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu }&\nu <M\\(g_{M}(x)-g_{M}^{*})/H_{M}&\nu =M\end{cases}}
+    \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu }&\nu <M\\(g_{M}(x)-g_{M}^{*})/H_{M}&\nu
+=M\end{cases}}
   \end{array}
@@
 
-Здесь \(M=\max_{}^{}\left\{\nu (x):x\in [0;1]\right\}\), а \(g_{M}^{*}=\min _{}^{}\left\{g_{M}(x):x\in \cap _{i=0}^{M-1}Q_{i}\right\}\).
-В силу определения числа \(M\), задача отыскания \(g_{M}^{*}\)
-всегда имеет решение, а если \(M=m+1\), то \(g_{M}^{*}=\varphi(x^{*})\).
-Дуги функции \(\psi (x)\) гельдеровы на множествах \(\cap _{i=0}^{j}Q_{i},0\leq j\leq M-1\)
-с константой 1, а сама \(\psi (x)\) может иметь разрывы первого рода на границах этих множеств.
-Несмотря на то, что значения констант Гёльдера \(H_k\) и величина \(g_{M}^{*}\) заранее неизвестны,
-они могут быть оценены в процессе решения задачи.
+Here \(M=\max_{}^{}\left\{\nu (x):x\in [0;1]\right\}\) and \(g_{M}^{*}=\min
+_{}^{}\left\{g_{M}(x):x\in \cap _{i=0}^{M-1}Q_{i}\right\}\).
+Because of the definition of the number \(M\), the problem of finding \(g_{M}^{*}\) always
+has a solution. If \(M=m+1\), then \(g_{M}^{*}=\varphi(x^{*})\).
+The arcs of the functions \(\psi (x)\) are the H\"{o}lder ones within the sets
+\(\cap _{i=0}^{j}Q_{i},0\leq j\leq M-1\)
+with the constant 1, and \(\psi (x)\) itself can have the breaks of the first kind at the boundaries
+of these sets.
+Despite the values of the H\"{o}lder constants \(H_k\) and the value \(g_{M}^{*}\) are not
+known in advance, these ones can be estimated in the course of solving the problem.
 
-Множество троек \(\{(x_k,\nu_k,z_k)\}, 1\leqslant k\leqslant n\) составляет поисковую информацию,
-накопленную методом после проведения \(n\) шагов.
+A set of triples \(\{(x_k,\nu_k,z_k)\}, 1\leqslant k\leqslant n\) constitutes the search information
+accumulated by the method upon executing \(n\) steps.
 
-На первой итерации метода испытание проводится в произвольной внутренней точке \(x_1\)
-интервала \([0;1]\). Индексы точек 0 и 1 считаются нулевыми, значения \(z\) в
-них не определены. Пусть выполнено \(k\geqslant 1\) итераций метода,
-в процессе которых были проведены испытания в \(k\) точках \(x_i, 1\leqslant i\leqslant k\).
-Тогда точка \(x^{k+1}\) поисковых испытаний следующей \((k+1)\)-ой
-итерации определяются в соответствии с правилами:
+At the first iteration of the method, the trial is performed in the arbitrary internal point \(x_1\) of
+the interval \([0;1]\).
+The indices of the points 0 and 1 are considered to be zero ones, the values \(z\) in these points
+are undefined.
+Let \(k\geqslant 1\) iterations of the method have been performed, in the course of which the
+trials were conducted in \(k\) points \(x_i, 1\leqslant i\leqslant k\).
+Then, the points \(x^{k+1}\) of the search trials of the next \((k+1)^{\rm th}\) iteration are
+defined according to the rules:
 
-Шаг 1. Перенумеровать точки множества \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\),
-которое включает в себя граничные точки интервала \([0;1]\), а также точки предшествующих
-испытаний, нижними индексами в порядке увеличения значений координаты, т.е.
-\begin{displaymath}
+Step 1. Renumber the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which
+includes the boundary points of the interval \([0;1]\) as well as the points of preceding trials by
+the lower indices in the order of increasing coordinate values i. e.
+\begin{equation}
+  \label{eq:points}
   0=x_0<x_1<\dotsc<x_{k+1}=1
-\end{displaymath}
-и сопоставить им значения \(z_{i}=g_{\nu }(x_{i}),\nu =\nu (x_{i}),i={\overline {1,k}}\).
+\end{equation}
+and juxtapose the values \(z_{i}=g_{\nu }(x_{i}),\nu =\nu (x_{i}),i={\overline {1,k}}\) to these
+ones.
 
-Шаг 2. Для каждого целого числа \(\nu ,1\leqslant \nu \leqslant m+1\) определить соответствующее
-ему множество \(I_{\nu }\) нижних индексов точек, в которых вычислялись значения
-функций \(g_{\nu }(x)\):
+Step 2. For each integer number \(\nu ,1\leqslant \nu \leqslant m+1\) determine the corresponding
+set \(I_{\nu }\) of the lower indices of the points, which the values of the functions \(g_{\nu
+}(x)\) are computed in:
 \begin{displaymath}
@@
 \end{displaymath}
-определить максимальное значение индекса \(M=\max\{\nu (x_{i}),1\leq i\leq k\}\).
+determine the maximum value of the index \(M=\max\{\nu (x_{i}),1\leq i\leq k\}\).
 
-Шаг 3. Вычислить текущие оценки для неизвестных констант Гёльдера:
+Step 3. Compute current estimate for the unknown H\"{o}lder constant:
 \begin{equation}
   \label{step2}
-  \mu _{\nu }=\max\{\frac{|g_{\nu }(x_{i})-g_{\nu }(x_{j})|}{(x_{i}-x_{j})^{\frac{1}{N}}}:i,j\in I_{\nu },i>j\}.
+  \mu _{\nu }=\max\{\frac{|g_{\nu }(x_{i})-g_{\nu }(x_{j})|}{(x_{i}-
+x_{j})^{\frac{1}{N}}}:i,j\in I_{\nu },i>j\}.
 \end{equation}
-Если множество \(I_{\nu }\) содержит менее двух элементов или если значение \(\mu _{\nu }\)
-оказывается равным нулю, то принять \(\mu _{\nu }=1\).
+If the set \(I_{\nu }\) contains less than two elements or if the value \(\mu _{\nu }\) appears to
+be equal to zero, then accept \(\mu _{\nu }=1\).
 
-Шаг 4. Для всех непустых множеств \(I_{\nu },\nu ={\overline {1,M}}\) вычислить оценки
-\begin{displaymath}
-  z_{\nu }^{*}={\begin{cases}\min\{g_{\nu }(x_{i}):x_{i}\in I_{\nu }\}&\nu =M\\-\varepsilon _{\nu }&\nu <M\end{cases}},
-\end{displaymath}
-где вектор с неотрицательными координатами \(\varepsilon _{R}=(\varepsilon _{1},..,\varepsilon _{m})\) называется вектором резервов.
+Step 4. For all nonempty sets \(I_{\nu },\nu ={\overline {1,M}}\) compute the estimates
+\begin{equation}
+  \label{eq:step_4}
+  z_{\nu }^{*}={\begin{cases}\min\{g_{\nu }(x_{i}):x_{i}\in I_{\nu }\}&\nu =M\\-\varepsilon
+_{\nu }&\nu <M\end{cases}},
+\end{equation}
+where the vector with the non-negative coordinates \(\varepsilon _{R}=(\varepsilon
+_{1},..,\varepsilon _{m})\) is called the vector of reserves.
 
-Шаг 5. Для каждого интервала \((x_{i-1};x_{i}),1\leqslant i\leqslant k\) вычислить характеристику
+Step 5. For each interval \((x_{i-1};x_{i}),1\leqslant i\leqslant k\) compute the characteristic
 \begin{equation}
   \label{step3_1}
-  R(i)={\begin{cases}\Delta _{i}+{\frac {(z_{i}-z_{i-1})^{2}}{(r_{\nu }\mu _{\nu })^{2}\Delta _{i}}}-2{\frac {z_{i}+z_{i-1}-2z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu (x_{i})=\nu (x_{i-1})\\2\Delta _{i}-4{\frac {z_{i-1}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu (x_{i-1})>\nu (x_{i})\\2\Delta _{i}-4{\frac {z_{i}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu (x_{i})>\nu (x_{i-1})\end{cases}}
+  R(i)={\begin{cases}\Delta _{i}+{\frac {(z_{i}-z_{i-1})^{2}}{(r_{\nu }\mu _{\nu
+})^{2}\Delta _{i}}}-2{\frac {z_{i}+z_{i-1}-2z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu
+(x_{i})=\nu (x_{i-1})\\2\Delta _{i}-4{\frac {z_{i-1}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu
+}}}&\nu =\nu (x_{i-1})>\nu (x_{i})\\2\Delta _{i}-4{\frac {z_{i}-z_{\nu }^{*}}{r_{\nu }\mu
+_{\nu }}}&\nu =\nu (x_{i})>\nu (x_{i-1})\end{cases}}
 \end{equation}
-где \(\Delta _{i}=(x_{i}-x_{i-1})^{\frac{1}{N}}\). Величины \(r_{\nu }>1,\nu ={\overline {1,m}}\)
-являются параметрами алгоритма. От них зависят произведения \(r_{\nu }\mu _{\nu }\),
-используемые при вычислении характеристик в качестве оценок неизвестных констант Гёльдера.
+where \(\Delta_{i}=(x_{i}-x_{i-1})^{\frac{1}{N}}\).
+The magnitudes \(r_{\nu }>1,\nu ={\overline {1,m}}\) are the parameters of the algorithm.
+The products \(r_{\nu}\mu_{\nu}\) used in computing the characteristics as the estimates of the
+unknown H\"{o}lder constants depend on these ones.
 
-Шаг 6. Выбрать наибольшую характеристику:
+Step 6. Select the maximum characteristic:
 \begin{equation}
@@
 
-Шаг 7. Провести очередное испытание в середине интервала \((x_{t-1};x_{t})\),
-если индексы его концевых точек не совпадают: \(x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})\).
-В противном случае провести испытание в точке
+Step 7. Perform the next trial in the middle of the interval \((x_{t-1};x_{t})\) if the indices of its
+end points are not the same: \(x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})\).
+Otherwise, perform the trial in the point
 \begin{displaymath}
-  x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})-\operatorname {sgn}(z_{t}-z_{t-1}){\frac {|z_{t}-z_{t-1}|^{n}}{2r_{\nu }\mu _{\nu }^{n}}},\nu =\nu (x_{t})=\nu (x_{t-1}),
+  x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})-\operatorname {sgn}(z_{t}-z_{t-1}){\frac {|z_{t}-
+z_{t-1}|^{n}}{2r_{\nu }\mu _{\nu }^{n}}},\nu =\nu (x_{t})=\nu (x_{t-1}),
 \end{displaymath}
-а затем увеличить \(k\) на 1.
+and then increment \(k\) by 1.
 
-Алгоритм прекращает работу, если выполняется условие \(\Delta_{t}\leqslant \varepsilon\),
-где \(\varepsilon>0\) есть заданная точность. В качестве оценки глобально-оптимального решения выбираются значения
+The algorithm stops if the condition \(\Delta_{t}\leqslant \varepsilon\) is satisfied where
+\(\varepsilon>0\) is a predefined precision.
+The value
 \begin{equation}
-\varphi_k^*=\min_{1\leqslant i \leqslant k}\varphi(x_i), x_k^*=\argmin_{1\leqslant i \leqslant k}\varphi(x_i)
+\varphi_k^*=\min_{1\leqslant i \leqslant k}\varphi(x_i), x_k^*=\argmin_{1\leqslant i \leqslant
+k}\varphi(x_i)
+\end{equation}
+is accepted as the estimate of the global optimal solution.
+
+Next, following the approach described in \cite{BarkalovStrongin2018}, we will use \(q\) copies
+of IAGS working synchronously to solve the problem series (\ref{eq:many_problems}).
+The only difference is that when selecting the interval with the maximum characteristic at Step
+6, the choice will be made from all intervals, which \(q\) copies of IAGS have generated to the
+moment.
+If the maximum characteristic corresponds to the \(i^{\rm th}\) problem, then Step 7 is
+performed in the \(i^{\rm th}\) copy of the method while the rest copies stay idle.
+This way, at every iteration the trial is performed in the most promising problem from the
+viewpoint of the characteristics (\ref{step3_1}) that allows distributing the resources of the
+method among the problems dynamically.
+
+The theory of convergence of such an approach for the case of solving the problems without
+constraints is presented in \cite{BarkalovStrongin2018}.
+If there are constraints, the characteristics of the intervals, at the boundaries of which different
+numbers of constraints are violated are computed according to the bottom rows of
+(\ref{step3_1}).
+One can notice that even in this case the magnitudes of characteristics are normalized as well.
+In the case of exact estimates \(z_{\nu }^{*}\) and \(\mu _{\nu }\), the scale of these ones
+doesn't depend on the objective function and constraints of particular problem.
+So far, the speculations from \cite{BarkalovStrongin2018} can be conducted in the case of the
+use of the index scheme as well.
+
+The parallel modification of the method doesn't differ from the one considered in
+\cite{BarkalovStrongin2018} and consists in the selection of \(p\) intervals at Step 6 and the
+performing of \(p\) trials in parallel at the next step.
+At that, all resources of the method within the framework of the iteration may be directed onto
+a single problem as well as onto \(l\leqslant p\) problems simultaneously (subject to the problem,
+which the intervals selected by the method belong to).
+
+\subsection{Convergence conditions }
+\label{sec:conv_method}
+
+The conditions of convergence of the described in Section \ref{sec:method} method in case of \(q=1\) are given in \cite{Strongin2000}.
+\begin{theorem} (Sufficient convergence conditions)
+  \label{th:single_conv}
+  Assume that the following conditions are true:
+  \begin{enumerate}
+    \item The problem (\ref{eq:constrained_problem}) has a solution.
+    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\) are Lipschitzian on \(D\) with
+respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
+    \item Beginning with some step (i. e. if \(k\) from (\ref{eq:points}) is sufficiently large),
+    the values \(\mu_\nu\) from \ref{step2} satisfy the inequalities
+    \begin{equation}
+      r_\nu\mu_\nu > 2^{3-1/N}L_\nu \sqrt{N+3},\: 1\leqslant \nu \leqslant m + 1.
+    \end{equation}
+  \end{enumerate}
+  Then any limit point \(y\) of the sequence \(\{y_k\} = \{y(x_k)\}\) generated
+  by the index method while solving the problerm
+  (\ref{eq:constrained_problem}) is admissible and satisfies the conditions
+\begin{equation}
+  \label{eq:conv_cond}
+  \varphi(\overline{y})=\{ \varphi(y^k): g_i(y^k)\leqslant 0,1\leqslant j\leqslant m, k=1,2,\dots\}=\varphi(y^*),
+\end{equation}
+  where \(y^*\) from (\ref{eq:constrained_problem}).
+\end{theorem}
+
+\begin{remark}
+  \label{rem:r1}
+  From the relationship between the Holder and Lipschitz constants and from the condition (\ref{eq:conv_cond}) it follows
+  that the parameters \(r_\nu\) from (\ref{step3_1}) should satisfy the condition
+  \begin{equation}
+    r_\nu > 2^{2 - 1/N}.
+  \end{equation}
+\end{remark}
+
+\begin{theorem} (On convergence condiotions) Let the conditions of the Theorem \ref{th:single_conv}
+  are satisfied for all of \(q\) problems from the set \ref{eq:many_problems} separately.
+  Then solving all of the \(q\) problems by the algorithm from Section \ref{sec:method} will
+  generate \(q\) infinite sequences, for the limit points of which the Theorem \ref{th:single_conv}
+  will be also true.
+\end{theorem}
+\begin{proof}
+  Let us consider two random problems from the set \ref{eq:many_problems}.
+  \begin{equation}
+      \begin{array}{lr}
+        \min\{\varphi(y):y\in D_1,\: g_j^\varphi(y)\leqslant 0, 1\leqslant j\leqslant m_1\} \\
+        \min\{\psi(y):y\in D_2,\: g_j^\psi(y)\leqslant 0, 1\leqslant j\leqslant m_2\}
+      \end{array}
+  \end{equation}
+  Let us denote the characteristics (\ref{step3_1}) for the first problem as \(R_\varphi(i)\)
+  and for the second problem as \(R_\psi(j)\). Considering that, we have:
+  \begin{equation}
+      \begin{array}{lr}
+        R_\varphi(t_\varphi)=\max_{1\leqslant i\leqslant k}R_\varphi(i), \\
+        R_\psi(t_\psi)=\max_{1\leqslant j\leqslant s}R_\psi(j),
+      \end{array}
+  \end{equation}
+  where \(k\) corresponds to the number of trials in the first problem and \(s\)
+  corresponds to the number of trials in the second one. The sequence of trials \(\{v^k\}\)
+  corresponds to the first problem and the sequence of trials \(\{u^s\}\)
+  corresponds to the second problem. The values \(z^k=g^\varphi_\nu(v_k),\nu =\nu (v_{k})\) correspond to the
+  trial points \(\{v^k\}\) and the values \(w^s=g^\psi_\nu(u_s),\nu =\nu(u_{s})\) correspond to the
+  trial points \(\{u^s\}\).
+
+  When the considered two problems are solved simultaneously, the algorithm selects
+  an interval for the next trial according to the condition:
+  \begin{equation}
+    R(t) = \max\{R_\varphi(i),R_\psi(j)\}
   \end{equation}
 
-Далее следуя подходу, описанному в \cite{BarkalovStrongin2018}, для решения серии задач (\ref{eq:many_problems}) будем
-использовать \(q\) синхронно работающих копий ИАГП с тем лишь отличием, что на шаге 6 при выборе
-интервала с наилучшей характеристикой, выбор будет осуществляться из всех интервалов, которые
-породили на данный момент \(q\) копий ИАГП. Если наибольшая характеристика соответствует
-задаче \(i\), то выполняется шаг 7 в копии метода с номером \(i\), а остальные копии метода простаивают.
-Таким образом, на каждой итерации испытание проводится в задаче, наиболее перспективной с точки зрения
-характеристик (\ref{step3_1}), что позволяет динамически распределять ресурсы метода между задачами.
+  Let the algorithm is solving tho problems and the iterations conunter is \(l = k + s, l=0,1,2,\dots\).
+  Then, since for all of the problems Theorem \ref{th:single_conv} is true, at least one of the sequences
+  \(\{v^k\}\) and \(\{u^s\}\) will be infinite (let it be \(\{v^k\}\)). If we prove that both the
+  sequences are infinite, this would lead us to convergence in both the considered problems.
 
-В \cite{BarkalovStrongin2018} приведена теория сходимости такого подхода на случай решения задач без ограничений.
-При наличии ограничений характеристики интервалов, на концах которых нарушено разное количество ограничений,
-вычисляются в соответствии с нижними строчками из (\ref{step3_1}). Нетрудно заметить, что и в этом случае
-величины характеристик нормированы, а в случае точных оценок \(z_{\nu }^{*}\) и \(\mu _{\nu }\) их
-масштаб не зависит от целевой функции и ограничений конкретной задачи. Таким образом, рассуждения
-из \cite{BarkalovStrongin2018} можно провести и в случае использования индексной схемы.
+  Let the limit point \(\overline{v}\in [v_{i-1},v_i]\), where \(i=i(k)\). The indexes of \(v_{i-1}\;,v_i\)
+  can be equal or different, but, because of convergence, starting from large enough \(k\) they will be stable.
+  %will be steady  equal to \(m_1+1\) (if \(\overline{v}\) is
+  %an interior point of the feasible domain produced by constraints \(g^\varphi_p(y),p=\overline{1,m_1}\))
+  %or \(\nu(v_{i-1})\ne\nu(v_i)\) (if \(\exists p: 1\leqslant p\leqslant m_1, g^\varphi_p(y(\overline{v}))=0\)).
+  In the first case the algorithm will use the first branch of the rule (\ref{step3_1}), otherwise it will use
+  one of the two other branches.
 
-Параллельная модификация метода не отличается от рассматриваемой в \cite{BarkalovStrongin2018}
-и заключается в выборе \(p\) интервалов на шаге 6 и выполнения \(p\) испытаний параллельно
-на следующем шаге. При этом все ресурсы метода в рамках итерации могут быть направлены как на одну, так и
-на \(l\leqslant p\) задач одновременно (в зависимости от того, какой из задач принадлежат выбранные методам интервалы).
+  Let us consider the first case: from (\ref{step2})
+  \begin{displaymath}
+    \frac{|z_i-z_{i-1}|}{\Delta_i} \leqslant \mu_{\varphi,\nu}.
+  \end{displaymath}
+  Considering that, we can establish an upper bound:
+  \begin{displaymath}
+    \frac{(z_i-z_{i-1})^2}{r_\nu^2\mu_{\varphi,\nu}^2\Delta_i}=\frac{(z_i-z_{i-1})^2\Delta_i}{(r_\nu\mu_{\varphi,\nu}\Delta_i)^2}
+    \leqslant \frac{\Delta_i}{r_\nu^2}.
+  \end{displaymath}
+  Therefore using the first branch form rule (\ref{step3_1}) we have inequality
+  \begin{equation}
+    \label{eq:th1}
+    R_\varphi(i)\leqslant\Delta_i(1 + \frac{1}{r_\nu^2}) - \frac{2(z_i+z_{i-1}-2z^*_\nu)}{r_\nu\mu_{\varphi,\nu}}
+  \end{equation}
+  Because \(\overline{v}\) is the limit point of the sequence \(\{v^k\}\) and \(\varphi(y(\overline{v}))\leqslant z^*_{\nu}, \nu=m+1\) or
+  \(z^*_\nu=0, \nu<m+1\) and \(z_{i-1},z_i\to 0\) at \(k\to\infty\):
+  \begin{equation}
+    \label{eq:th2}
+    \Delta_i\to 0, z_i+z_{i-1} - 2 z_\nu^*\to 0.
+  \end{equation}
+  In the second case (when one of the other two branches rule (\ref{step3_1}) is applied) from we have
+  \begin{displaymath}
+    R_\varphi(i)=2\Delta_i + 4\frac{z^*_\nu-z_i}{r_\nu\mu_{\varphi,\nu}}
+  \end{displaymath}
+  If \(z^*_\nu \ne 0\) then \(z_i-z^*_\nu \geqslant 0\) and
+  \begin{equation}
+    \label{eq:th3}
+    R_\varphi(i)=2\Delta_i - 4\frac{z_i-z^*_\nu}{r_\nu\mu_{\varphi,\nu}} \leqslant 2\Delta_i.
+  \end{equation}
+  Otherwise because of \(\overline{v}\) is feasible point \(z_i\to 0\) at \(k\to\infty\).
 
-\section{Результаты численных экспериментов}
+  From (\ref{eq:th1}) (\ref{eq:th2}) (\ref{eq:th3}) for any small \(\delta > 0\) exists a large value of \(k\) such that
+  \begin{equation}
+    R_\varphi(i)\leqslant \delta.
+    \label{eq:th6}
+  \end{equation}
+
+  Let \(\alpha = \max\{\nu(u^s)\}\). Because \(\alpha\) is currently the greatest index
+  in the search sequence \(\{u^s\}\) and according to the rule (\(\ref{eq:step_4}\)), \(\exists j: w^*_\alpha=w_j\).
+
+  If \(\nu(w_{j-1})=\nu(w_{j})=\alpha\) then the first branch of the rule (\ref{step3_1}) is applied and
+  \begin{dmath}
+    R_\psi(j)=\Delta_j + \frac{(w_j-w_{j-1})^2}{r_\alpha^2\mu_{\psi,\alpha}^2\Delta_j}
+      - \frac{2(w_j+w_{j-1}-2w^*_\alpha)}{r_\alpha\mu_{\psi,\alpha}} \geqslant
+      \Delta_j - \frac{2(w_j+w_{j-1}-2w^*_\alpha)}{r_\alpha\mu_{\psi,\alpha}} =
+      \Delta_j - \frac{2\Delta_j(w_{j-1}-w_j)}{r_\alpha\mu_{\psi,\alpha}\Delta_j} \geqslant
+      \Delta_j - \frac{\Delta_j}{r_\alpha} = \Delta_j\left(1-\frac{2}{r_\alpha}\right).
+    \label{eq:th4}
+  \end{dmath}
+
+  If \(\nu(w_{j-1})\ne\nu(w_{j})=\alpha\) then \(\nu(w_{j-1})<\alpha\) and the third branch of
+  the rule (\ref{step3_1}) is applied:
+  \begin{equation}
+    R_\psi(j)=2\Delta_j - 4 \frac{w_j-w^*_{\alpha}}{r_\alpha \mu_{\psi,\alpha}}=2\Delta_j > 0.
+    \label{eq:th5}
+  \end{equation}
+  Taking into account the Remark \ref{rem:r1}, (\ref{eq:th4}), (\ref{eq:th5}) we can conclude that
+  exists such interval that \(R_\psi(j)>0\). At the same time, (\ref{eq:th6}) is true and
+  the inequality
+  \begin{displaymath}
+    R_\psi(j) > R_\varphi(i)
+  \end{displaymath}
+  will be true at large enough value of \(k\). Thus the next scheduled iteration is performed on
+  the second problem with objective \(\psi(y)\), i.e. sequence \(\{v^s\}\) will be infinite as well.
+
+  Since we considered two arbitrary problems from the given set of \(q\) problems, the theorem is
+  true for any pair of problems from the set. By induction the theorem is also true for the whole
+  set.
+\qed
+\end{proof}
+
+\section{Results of numerical experiments}
 \label{sec:exps}
 
-Использование сгенерированных некоторыми случайными механизмами
-наборов тестовых задач с известными решениями является одним из общепринятых подходов
-к сравнению алгоритмов оптимизации \cite{Beiranvand2017}. В данной работе
-будем использовать два генератора тестовых задач, порождающих задачи различной природы \cite{grishaginClass, Gaviano2003}.
-Эти генераторы порождают задачи без нелинейных ограничений, поэтому в дополнение к ним использована
-система GCGen\footnote{Исходный код системы доступен по ссылке https://github.com/UNN-ITMM-Software/GCGen} \cite{GergelBarkalov2019}, позволяющая генерировать задачи с ограничениями на основе произвольных нелинейных
-функций.
+The use of set of test problems with known solutions generated by some random mechanisms is
+one of commonly accepted approaches to the comparing of the optimization algorithms
+\cite{Beiranvand2017}.
+In the present work, two generators of test problems generating the problems of different nature
+\cite{grishaginClass, Gaviano2003} were used.
+These generators generate the problems without the nonlinear constraints. Therefore, the
+GCGen\footnote{The original code of the system can be accessed at https://github.com/UNN-
+ITMM-Software/GCGen} \cite{GergelBarkalov2019} system was used in addition.
+This system allows generating the problems with constraints based on arbitrary nonlinear
+functions.
 
-Вместе с системой GCGen распространяются примеры ее использования и построения
-наборов задач, каждая из которых состоит из целевой функции и двух ограничений,
-порожденных генератором \(F_{GR}\) \cite{grishaginClass} или GKLS \cite{Gaviano2003}.
-
-% Обозначим набор из 50 задач, полученных с помощью GCGen и первого генератора из \cite{grishaginClass} как \(F_{GR}^C\).
-Механизм построения функций \(F_{GR}\) не предусматривает контроля за размерностью (она равна 2) и количеством локальных оптимумов, однако известно, что порождаемые функции
-% являются существенно многоэкстремальными.
-
-Генератор GKLS \cite{Gaviano2003} позволяет получать функции заданной размерности и с заданным количеством экстремумов.
-В сочетании с GCGen были порождены два множества по 100 задач размерности 2 и 3. Каждая из задач имеет два ограничения.
-Также с целью демонстрации того, что свойства метода сохраняются при существенно разных свойствах задач
-был сгенерирован смешанный класс, состоящий из 50 задач с двухмерными функциями GKLS и 50 задач с функциями \(F_{GR}\).
-На рис. \ref{fig:isolines} представлены примеры линий уровня рассматриваемых задач. Допустимая область закрашена.
+Along with GCGen system, the examples of its application and constructing the set of problems
+each consisting of an objective function and two constraints generated by \(F_{GR}\)
+\cite{grishaginClass} generator or GKLS \cite{Gaviano2003} one Are distributed.
+GKLS \cite{Gaviano2003} allows obtaining the functions of predefined dimensionality and
+with predefined number of extrema.
+In combination with GCGen, two sets of 100 problems each of the dimensionalities 2 and 3
+were generated.
+Each problem had two constraints.
+Also, in order to demonstrate the properties of the method to be preserved at essentially
+different properties of the problems, a mixed class consisting of 50 problems with two-
+dimensional GKLS functions and 50 problems with \(F_{GR}\) ones was generated.
+Examples of contour plots of the considered problems are presented in Fig. \ref{fig:isolines}.
+The feasible area is highlighted by color.
 
@@
     \centering
-    \subfloat[Решение внутри допустимой области]{
+    \subfloat[Solution inside feasible domain]{
     {\includegraphics[width=.5\textwidth]{pic/2.png}}}
-    \subfloat[Решение на границе допустимой области]{
+    \subfloat[Solution at the boundary of feasible area]{
     {\includegraphics[width=.5\textwidth]{pic/4.png}}}
-    \caption{Линии уровня и точки испытаний ИАГП в двух синтетических задачах}
+    \caption{Contour plots and trial points of IAGS in two synthetic problems}
     \label{fig:isolines}
@@
 
-Будем считать, что тестовая задача решена, если метод оптимизации провел очередное испытание \(y^k\) в
-\(\delta\)-окрестности глобального минимума \(y^*\), т.е. $\left\|y^k-
-y^*\right\|\leqslant \delta = 0.01\left\|b-a\right\|$, где \(a\) и \(b\) --- левая и правая границы гиперкуба из (\ref{eq:task}).
-Если указанное соотношение не выполнено до истечения лимита на количество испытаний, то задача считается нерешенной.
+A test problem was considered to be solved if the optimization method executes the next trial
+\(y^k\) in the \(\delta\)-proximity of the global minimum \(y^*\) i. e.
+\(\left\|y^k-y^*\right\|\leqslant \delta = 0.01\left\|b-a\right\|\)
+Where \(a\) and \(b\) are the left and right boundaries of the hypercube from (\ref{eq:task}).
+If this relation is not fulfilled upon reaching the limit of number of iterations, the problem was
+considered to be not solved.
 
-%В качестве характеристик метода оптимизации на каждом из классов будем рассматривать среднее число
-%испытаний, затраченное для решения одной задачи, и количество решенных задач. Чем меньше число испытаний, тем быстрее метод сходится
-%к решению, а значит и меньше обращается к потенциально трудоемкой процедуре вычислений целевой функции.
-При оценке качества метода и его реализации кроме ускорения от распараллеливания и времени выполнения также будем принимать во внимание среднее максимальное расстояния (в смысле \(l_{\inf}\)-нормы) текущей оценки оптимума до его реального положения,
-вычисленное на множестве задач (\ref{eq:many_problems}): \(D_{avg}\) и \(D_{max}\). Динамика этих величин в процессе оптимизации
-показывает, насколько равномерно метод распределяет ресурсы между задачами.
+When evaluating the quality of the method and its implementation, besides the speedup due to
+the parallelization and the work time,
+Was taken into account as well.
+The mean maximum distance (in terms of \(l_{\inf}\)-norm) form the current estimate of the
+optimum to its actual position computed on the set of problems (\ref{eq:many_problems}):
+\(D_{avg}\) and \(D_{max}\).
+The dynamics of these magnitudes in the course of the optimization shows how uniformly the
+method distributes the resources among the problems.
 
-Реализация параллельного метода была выполнена на языке C++ с использованием технологии OpenMP
-для распареллеливания процесса проведения испытаний на общей памяти. Все вычислительные
-эксперименты проведены на машине со следующей конфигурацией: Intel Core i7-7800X, 64GB RAM, Unubtu 16.04 ОS, GCC 5.5 compiler.
+The parallel method was implemented on C++ language with the use OpenMP technology for
+the parallelization of the trial execution process in the shared memory.
+All numerical experiments were carried out using a computer with the following  configuration:
+Intel Core i7-7800X, 64GB RAM, Ubuntu 16.04 OS, GCC 5.5 compiler.
 
-\subsection{Результаты решения сгенерированных задач}
+\subsection{Result of solving the generated problems}
 
-Результаты решения тестовых задач последовательной и параллельной версией модифицированного ИАГП
-для решения множества задач представлены в \ref{tab:speedup}. Для всех двухмерных классов задач параметр \(r=4.7\).
-В случае трехмерных задач \(r=4.7,\: \varepsilon_\nu=0.1\).
-Во всех экспериментах в целевые функции и ограничения была внесена дополнительная вычислительная нагрузка так,
-чтобы время одного обращения к функции задачи было равно примерно 1мс.
+The results of solving the test problems by the sequential and parallel versions of the modified
+IAGS for solving the sets of problems are presented in Table \ref{tab:speedup}.
+For all two-dimensional classes of problems, the parameter \(r=4.7\).
+In the case of the three-dimensional problems, \(r=4.7,\: \varepsilon_\nu=0.1\).
+In all experiments, an additional computational load was introduced into the objective function
+and constraints so that the time of a single call of a problem function was equal to
+approximately 1 ms.
 
-Из таблицы видно, что ускорение по итерациям \(S_i\) растет линейно с увеличением числа потоков \(p\),
-в то время, как ускорение по времени \(S_p\) увеличивается не так быстро, что говорит о неидеальной
-реализации алгоритма. Увеличить реальное ускорение, верхней границей для которого является
-\(S_i\), возможно путем оптимизации взаимодействий между копиями ИАГП и это планируется сделать в ходе будущей работы.
+One can see from Table \ref{tab:speedup} that the speedup in the iterations \(S_i\) increased
+linearly with increasing number of threads \(p\) whereas the speedup in time \(S_t\) increased
+not so fast that points to a non-ideal implementation of the algorithm.
+One can increase the actual speedup, the upper limit for which is \(S_i\) by the optimization of
+the interaction between the copies of IAGS. It is planned to do in the course of future work.
 
@@
   \centering
-  \caption{Результаты экспериментов на наборах синтетических задач}
+  \caption{Results of experiments on the sets of synthetic problems}
   \label{tab:speedup}
@@
     %\cline{1-8}\noalign{\smallskip}
-    Класс задач & \textit{p} & Количество итераций & Время, с & \(S_i\) & \(S_t\)   \\
+    Problem class & \textit{p} & Number of iterations & Time, s & \(S_i\) & \(S_t\)   \\
     %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
@@
 
-Для того, чтобы показать равномерную сходимость все тестовые задачи были также решены ИАГП
-в режиме решения отдельных задач. На рис. \ref{fig:devs_mixed} указаны графики величин
-средних и максимальных расстояний от реальных оптимумов до их текущих оценок при решении
-серии из задач, порожденных двумя разными генераторами, по отдельности (сплошная кривая) и совместно (пунктирная кривая).
-Не смотря на значительную разницу в структуре задач, модифицированный ИАГП
-гораздо быстрее уменьшает как максимальное, так и среднее отклонение оценок от оптимумов.
-Это говорит о наличии равномерной сходимости по всему множеству совместно решаемых задач.
-При этом в случае последовательного решения задач величина \(D_{max}\) имеет наибольшее значение вплоть
-до решения последней задачи.
+In order to demonstrate the uniform convergence, all test problems have been solved by IAGS
+in the regime of solving separate problems as well.
+Fig. \ref{fig:devs_mixed} presents the plots of mean and maximum distances from the actual
+optima to the current estimates of these ones when solving a series of problems generated by
+two different generators separately (solid line) and together (dashed line).
+In spite of an essential differences in the structure of problems, the modified IAGS decreased
+the maximum deviations of the estimates from the optima as well as the mean ones much faster.
+It evidences the uniform convergence over the whole set of problems being solved all together.
+At that, in the case of the sequential solving of the problems, the magnitude \(D_{max}\) takes
+its maximum value until the last problem is solved.
 
@@
     \label{fig:avg_dev} {\includegraphics[width=.5\textwidth]{pic/mixed_2d_avg.pdf}}}
-    \caption{Динамика величин \(D_{avg}\) и \(D_{max}\) в процессе решения множества двухмерных задач,
-    порождённых двумя разными генераторами GKLS и \(F_{GR}\)}
+    \caption{Dynamics of \(D_{avg}\) and \(D_{max}\) in the course of solving a set of the two-
+dimensional problems generated by two different generators GKLS and \(F_{GR}\)}
     \label{fig:devs_mixed}
@@
 
-\subsection{Пример решения многокритериальной задачи}
+\subsection{Example of solving a multicriteria problems}
 
-Для демонстрации эффективности подхода к балансировке нагрузки рассмотрим пример,
-в котором множество задач вида (\ref{eq:many_problems}) порождено в результате скаляризации
-многокритериальной задачи оптимизации с ограничениями.
+In order to demonstrate the efficiency of the approach to the balancing of the load, let us
+consider an example, in which a set of problems of the kind (\ref{eq:many_problems}) is
+generated as a result of scalarization of a multicriteria optimization problem with constrains.
 
-Рассмотрим тестовую задачу, предложенную в \cite{BinhKorn1999}:
+Let us consider a test problem proposed in \cite{BinhKorn1999}:
 \begin{equation}
@@
 
-Будем использовать свертку Гермейера для скаляризации задачи (\ref{eq:mco_probem}).
-После свертки скалярная целевая функция имеет вид:
+Let us use the Germeyer convolution for the scalarization of the problem
+(\ref{eq:mco_probem}).
+After the convolution, the scalar objective function takes the form:
 \begin{equation}
@@
 \end{equation}
-где \(\lambda_1,\lambda_2\in[0,1],\: \lambda_1+\lambda_2=1\). Перебирая все возможные
-коэффициенты свертки, можно найти все множество парето-оптимальных решений в
-задаче (\ref{eq:mco_probem}). Для численного построения множества Парето выберем
-100 наборов коэффициентов \((\lambda_1,\lambda_2)\) таких, что
+where \(\lambda_1,\lambda_2\in[0,1],\: \lambda_1+\lambda_2=1\).
+Trying all possible convolution coefficients, one can find the whole set of Pareto-optimal
+solutions in the problem (\ref{eq:mco_probem}).
+For the numerical construction of the Pareto set, let us select 100 sets of coefficients
+\((\lambda_1,\lambda_2)\) so that
 \(\lambda_1^i=i h,\: \lambda_2^i=1-\lambda_1^i,\: h=10^{-2},i=\overline{1, 100}\).
 
-В качестве ограничения на вычислительные ресурсы был выбран лимит в 2500 испытаний.
-Множество вспомогательных скалярных задач решалось двумя способами:
+As the limitation for the computational resources, a limit of 2500 trials was selected.
+The set of auxiliary scalar problems was solved by two methods:
 \begin{itemize}
-  \item каждая задача решается отдельно с помощью ИАГП с установленным лимитом в
-  25 испытаний. Таким образом, вычислительные ресурсы равномерно распределены между задачами;
-  \item все задачи решаются одновременно с помощью обобщенного ИАГП с установленным лимитом в
-  2500 испытаний.
+  \item each problem was solved separately using IAGS with a preset limit of 25 trials. This way,
+the computational resources were distributed among the problems uniformly;
+  \item all problems are solved simultaneously using generalized IAGS with a preset limit of
+2500 trials.
 \end{itemize}
-В обоих случаях параметр \(r=4\).
+In both cases, the parameter \(r=4\).
 
@@
     \centering
-    \subfloat[ИАГП, раздельное решение задач]{
+    \subfloat[IAGS, separate solving of the problems]{
     \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{pic/single_mco.pdf}}}
-    \subfloat[ИАГП для множества задач]{
+    \subfloat[IAGS for the set of problems]{
     \label{fig:mco_pareto_2} {\includegraphics[width=.5\textwidth]{pic/multi_mco.pdf}}}
-    \caption{Численные оценки множества Парето в задаче (\ref{eq:mco_probem}), полученные после 2500 испытаний}
+    \caption{Numerical estimates of Pareto set in the problem (\ref{eq:mco_probem}), obtained
+after 2500 trials}
     \label{fig:mco_pareto}
@@
 
-На рис. \ref{fig:mco_pareto_1} и рис. \ref{fig:mco_pareto_2}
-представлены графики решений, полученных каждым из методов.
-Все графики качественно совпадают с указанным в \cite{BinhKorn1999} (авторы не предоставили другой
-информации и решениях для сравнения). Видно, что на рис. \ref{fig:mco_pareto_1}
-кривая Парето имеет вогнутости, что не соответствует решению, указанному в \cite{BinhKorn1999} и
-означает нехватку ресурсов для решения некоторых из вспомогательных задач.
-Для оценки качества решения также был вычислен показатель \(Spacing(SP)\) \cite{RiquelmeLucken2015},
-характеризующий плотность точек аппроксимации множества Парето.
+The plots of solutions obtained by each method are presented in Fig. \ref{fig:mco_pareto_1}
+and Fig. \ref{fig:mco_pareto_2}.
+All plots agree with the ones presented in \cite{BinhKorn1999} qualitatively (the authors
+haven't provided any other information to compare).
+One can see the Pareto curve in Fig. \ref{fig:mco_pareto_1} to have concavities that doesn't
+match the solution presented in \cite{BinhKorn1999} and means the lack of resources for
+solving some auxiliary problems.
+To evaluate the quality of solution, the index \(Spacing(SP)\) \cite{RiquelmeLucken2015}
+featuring the density of the points approximating the Pareto set was computed also.
 \begin{displaymath}
@@
 \end{displaymath}
-В случае отдельного решения задач \(SP_{single}=0.984\), а при решении задач методом с балансировкой нагрузки
-\(SP_{multi}=0.749\), что говорит о более качественном приближении решения.
+In the case of separate solving of the problem \(SP_{single}=0.984\) whereas when solving of
+the problems by the method with the balancing of the load \(SP_{multi}=0.749\) that evidences
+a better quality of the approximation of the solution.
 
-\section{Заключение}
+\section{Conclusion}
 
-В ходе работы была реализована поддержка нелинейных ограничений в алгоритме, решающeм
-множество задач глобальной оптимизации. Проведены численные эксперименты, демонстрирующие
-преимущество рассматриваемого подхода над решением задач по отдельности. Показана эффективность
-совместного решения множества задач на примере решения многокритериальной задачи с
-нелинейными ограничениями.
+In the course of present work, the support of the nonlinear constraints was implemented in the
+algorithm solving a set of the global optimization problems.
+Also, the essential conditions of convergence for the developed method were obtained.
+The numerical experiments demonstrating the advantages of the considered approach over
+solving the problems separately were carried out.
+The efficiency of solving a set of problems together was demonstrated by the example of
+solving a multicriteria problem with nonlinear constraints.
 
-В ходе дальнейшей работы планируется улучшить текущую реализацию алгоритма,
-сократив расходы на содержание поисковой информации для множества задач и тем самым улучшив
-показатели параллельного ускорения по времени. Также планируется реализовать версию
-рассматриваемого алгоритма, работающего на распределенной памяти по схеме, описанной в \cite{BarkalovLebedev2017_2}.
+In the course of further work, it is planned to improve current implementation of the algorithm
+by reducing the costs of supporting the search information for the set of problems and, this way,
+improving the indices of speedup in time due to the parallelization.
+Also, it is planned to implement a version of considered algorithm working in the distributed
+memory according to the scheme described in \cite{BarkalovLebedev2017_2}.
 
@@
 %
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
+\bibliographystyle{splncs04}
+\bibliography{bibliography}
 %
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
 \end{document}

### 2020-06-13 2020/OPTIMA/optima_paper.tex
Check Section 1 and 2

@@
 %
-\title{Parallel global optimization algorithm for obtaining uniform convergence when
-simultaneously solving a set of global optimization problems\thanks{This study was supported
+\title{Parallel global optimization algorithm with uniform convergence
+for solving a set of constrained global optimization problems\thanks{This study was supported
 by the Russian Science Foundation, project No.\,16-11-10150.}}
@@
 %
-\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
+\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
 \email{sovrasov.vladislav@itmm.unn.ru},
@@
 \begin{abstract}
-In the present work, an extension of global optimization algorithm solving a number of problems
-simultaneously onto a case of problems with non-linear constraints is considered. The property
+In the present work, an extension of global optimization algorithm for solving a set of problems
+simultaneously onto a case of problems with non-convex constraints is considered. The property
 of uniform convergence inherent to the algorithm allows distributing the computation resources
@@
 example, in solving a multicriteria optimization problem using criteria convolution. The
-considered algorithm employs mappings like Peano space filling curve for the reduction of the
+considered algorithm employs Peano-type space-filling curves for the reduction of the
 multidimensional optimization problems to the one-dimensional ones.  The efficiency of the
@@
 multicriteria optimization  problem. Also, the numerical experiments confirmed the uniform
-convergence of the method.
+convergence of the proposed method.
 
-\keywords{Global optimization \and Uniform convergence \and Derivative-free optimization
-\and Parallel computations.}
+\keywords{Global optimization \and Non-convex constraints \and Uniform convergence \and Derivative-free optimization
+\and Parallel computing.}
 \end{abstract}
@@
 \label{sec-intro}
-Nonlinear global optimization of the non-convex functions is considered traditionally to be one
-of the most difficult problems of mathematical programming.
+Global optimization problems with non-convex constraints are considered traditionally to be one
+of the most difficult optimization problems.
  The finding of the global minimum of a function of several variables appears to be more
 difficult often than a local optimization in a thousand-dimensional space.
-For the latter, application of the simplest gradient descent method may appear to be enough
-whereas in order to \textit{guarantee} the finding of the global optimum, the optimization
-methods have to accumulate the information on the behavior of the objective function in the
+For the latter, application of the simplest gradient descent method may appear to be enough,
+whereas to \textit{guarantee} the finding of the global optimum the optimization
+methods have to accumulate the information on the behaviour of the objective function in the
 whole search domain \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}.
@@
 In the present work, a generalization of parallel global optimization method for simultaneous
-solving of a set of problems developed earlier in Lobachevskii University of Nizhnii Novgorod
-\cite{BarkalovStrongin2018} onto the case of nonlinear constraints is considered. To account
-for the constraints, the index scheme \cite{Strongin2000} is applied allowing working with
-partially computable objective functions and featured by efficiency comparable to other
+solving of a set of problems developed earlier in Lobachevsky State University of Nizhni Novgorod
+\cite{BarkalovStrongin2018} onto the case of non-convex constraints is considered. To account
+for the constraints, the index scheme \cite{Strongin2000,Pugliese} is applied allowing working with
+partially defined objective function and constraints and featured by efficiency comparable to other
 approaches \cite{BarkalovLebedev2017}. The convergence of the algorithm to the global
-optimal solutions in all problems was proved theoretically, the efficiency of the implemented
+optimizers in all problems was proved theoretically, the efficiency of the implemented
 algorithm was demonstrated by solving sets of problems generated by a special mechanism
 generating the sets of problems of predefined dimensionality with predefined number of
-nonlinear constraints \cite{GergelBarkalov2019}.
+non-convex constraints \cite{GergelBarkalov2019}.
 Besides the problems generated artificially, the considered method was tested also on a set of
@@
 \label{sec:problem}
-Within the frames of the present work, we will consider the following global optimization
-problem statement: to find the global minimum of an \(N\)-dimensional function \(\varphi(y)\) in
-a hyperinterval
-\(D=\{y\in \mathbf{R}^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
+In this work we consider the following 
+problem: to find the global minimum of an \(N\)-dimensional function \(\varphi(y)\) in
+a hyperinterval $D$
+\begin{equation}
+\label{eq:task}
+\varphi(y^*)=\min\{\varphi(y):y\in D\},
+\end{equation}
+\begin{displaymath}
+D=\{y\in \mathbf{R}^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}.
+\end{displaymath}
 In order to construct an estimate of the global minimum based on a finite number of
@@
 be limited.
-As a rule, the Lipschitz condition Is accepted as such a limitation.
-\begin{equation}
-\label{eq:task}
-\varphi(y^*)=\min\{\varphi(y):y\in D\}
-\end{equation}
+As a rule, the Lipschitz condition 
 \begin{displaymath}
 \label{lip}
-|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
+|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty,
 \end{displaymath}
+is accepted as such a limitation.
 
 There are various methods solving the considered multidimensional problem directly
-\cite{SergeyevKvasov2017, Jones2009} as well as efficient methods of solving the one-
-dimensional problems \cite{Norkin1992, Strongin2000}.
+\cite{SergeyevKvasov2017, Jones2009} as well as efficient methods for solving the univariate problems \cite{Norkin1992, Strongin2000}.
 In the present work, a one-dimensional method is considered, which is applied together with the
 dimensionality reduction scheme.
-The use of evolvents or space-filling curves is a classical scheme of reduction of the
-dimensionality of the initial problem for the global optimization algorithms
-\cite{Sergeyev2013}.
+The use of space-filling curve (or \textit{evolvent}) $y(x)$, where
 \begin{equation}
@@
 \lbrace y\in \mathbf{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant
-N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
+N\rbrace=\{y(x):0\leqslant x\leqslant 1\},
 \end{equation}
-
-A mapping of the kind (\ref{cube}) allows reducing a problem in the multidimensional space to
-solving a one-dimensional one at the expense of worsening its properties. In particular, the one-
-dimensional function \(\varphi(y(x))\) is not a Lipschitz but a H\"{o}lder one:
+ is a classical scheme of reduction of the
+dimensionality of the initial problem for the global optimization algorithms
+\cite{Sergeyev2013}.
+A mapping of the type (\ref{cube}) allows reducing a multidimensional problem to a univariate one at the expense of worsening its properties. In particular, the 
+one-dimensional function \(\varphi(y(x))\) is not lipschitzian but h\"{o}lderian one:
 \begin{displaymath}
 \label{holder}
-|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0;1],
+|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}}, \; x_1,x_2\in[0,1],
 \end{displaymath}
@@
 
-The domain \(D\) can also be defined by the constraint functions that complicates the problem
-essentially.
-The statement of the global optimization problem in this case will take the following form:
+The feasible domain \(G\) can also be defined by the constraints that essentially complicate the problem.
+The problem statement in this case will take the following form:
 \begin{equation}
   \label{eq:constrained_problem}
-  \varphi(y^*)=\min\{\varphi(y):y\in G=D \cap \{y: g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}\}
+  \varphi(y^*)=\min\{\varphi(y):y\in G=D \cap \{y: g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}\}.
 \end{equation}
@@
 Let us denote \(g_{m+1}(y)=\varphi(y)\). Hereafter, we will assume all functions
-\(g_k(y),1\leqslant k \leqslant m+1\)
+\(g_k(y),1\leqslant k \leqslant m+1\),
 to satisfy the Lipschitz condition in some hyperinterval including \(D\).
@@
 describing the method that it is required to find the global minimum of the function \(\varphi(x),
-x\in[0;1]\) satisfying the H\"{o}lder condition with the constraints \(g_j(x)\) also satisfying this
-condition in the interval \([0;1]\).
+x\in[0,1]\), satisfying the H\"{o}lder condition with the constraints \(g_j(x)\) also satisfying this
+condition in the interval \([0,1]\).
 
 The index algorithm of global search (IAGS) for solving the one-dimensional problems
-(\ref{eq:constrained_problem}) considered here implies the constructing of a sequence of points
+%(\ref{eq:constrained_problem}) 
+considered here implies the constructing of a sequence of points
 \(x_k\), which the values of the minimized function or of the constraints \(z_k = g_s(x_k)\) are
@@
 To account for the latter, the index scheme \cite{Strongin2000} was used.
-Let \(Q_0=[0;1]\). The constraint having the index \(j\) is satisfied in all points of the domain
+Let \(Q_0=[0,1]\). The constraint having the index \(j\) is satisfied in all points of the domain
 \begin{displaymath}
-  Q_j=\left\{x\in [0;1]:g_j(x)\leq 0\right\},
+  Q_j=\left\{x\in [0,1]:g_j(x)\leq 0\right\},
 \end{displaymath}
 which is called feasible for this constraint.
-At that, the feasible domain \(D\) of the initial problem is defined by the equality:
+At that, the feasible domain \(D\) of the initial problem is defined by the equality
  \(D=\cap _{j=0}^{m}Q_{j}\).
-A trial in the point \(x\in [0;1]\) consists in a sequence computing of the values
+A trial in the point \(x\in [0,1]\) consists in a sequence computing of the values
 \(g_{1}(x),...,g_{\nu }(x)\)
@@
 \(x\in Q_{j},0\leqslant j<\nu ,x\notin Q_{\nu }\).
-The finding of the first violation of any constraint terminates the trial in the point \(x\).
-In the case when the point \(x\) is feasible i. e. \(x\in D\), the trial includes the computing of all
+The finding of the first violation of any constraint terminates the trial at the point \(x\).
+In the case when the point \(x\) is feasible, i.e. \(x\in D\), the trial includes the computing of all
 functions of the problem.
@@
 The pair \(\nu =\nu (x), z=g_{\nu }(x)\) where the index \(\nu\) falls into the borders \(1\leqslant
-\nu \leqslant m+1\) is called the result of the trial in the point \(x\).
+\nu \leqslant m+1\) is called the result of the trial at the point \(x\).
 
@@
   \begin{array}{lr}
-    \psi (x^{*})=\min_{x\in [0;1]}\psi (x), \\
-    \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu }&\nu <M\\(g_{M}(x)-g_{M}^{*})/H_{M}&\nu
+    \psi (x^{*})=\min_{x\in [0,1]}\psi (x), \\
+    \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu },&\nu <M\\(g_{M}(x)-g_{M}^{*})/H_{M},&\nu
 =M\end{cases}}
@@
 
-Here \(M=\max_{}^{}\left\{\nu (x):x\in [0;1]\right\}\) and \(g_{M}^{*}=\min
+Here \(M=\max_{}^{}\left\{\nu (x):x\in [0,1]\right\}\) and \(g_{M}^{*}=\min
 _{}^{}\left\{g_{M}(x):x\in \cap _{i=0}^{M-1}Q_{i}\right\}\).
@@
 At the first iteration of the method, the trial is performed in the arbitrary internal point \(x_1\) of
-the interval \([0;1]\).
+the interval \([0,1]\).
 The indices of the points 0 and 1 are considered to be zero ones, the values \(z\) in these points
@@
 Step 1. Renumber the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which
-includes the boundary points of the interval \([0;1]\) as well as the points of preceding trials by
+includes the boundary points of the interval \([0,1]\) as well as the points of preceding trials by
 the lower indices in the order of increasing coordinate values i. e.
@@
 end points are not the same: \(x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})\).
-Otherwise, perform the trial in the point
+Otherwise, perform the trial at the point
 \begin{displaymath}

### 2020-06-13 2020/OPTIMA/optima_paper.tex
OPTIMA: fix some typos in the proof

@@
   \end{equation}
-  In the second case (when one of the other two branches rule (\ref{step3_1}) is applied) from we have
+  In the second case (when one of the other two branches of the rule (\ref{step3_1}) is applied) we have
   \begin{displaymath}
-    R_\varphi(i)=2\Delta_i + 4\frac{z^*_\nu-z_i}{r_\nu\mu_{\varphi,\nu}}
+    R_\varphi(i)=2\Delta_i - 4\frac{z_i-z^*_\nu}{r_\nu\mu_{\varphi,\nu}}
   \end{displaymath}
@@
   \end{equation}
-  Otherwise because of \(\overline{v}\) is feasible point \(z_i\to 0\) at \(k\to\infty\).
+  Otherwise because of \(\overline{v}\) is a feasible point \(z_i\to 0\) at \(k\to\infty\).
 
@@
 
-  Let \(\alpha = \max\{\nu(u^s)\}\). Because \(\alpha\) is currently the greatest index
+  Let \(\alpha = \max\{\nu(u):u\in\{u^s\}\}\). Because \(\alpha\) is currently the greatest index
   in the search sequence \(\{u^s\}\) and according to the rule (\(\ref{eq:step_4}\)), \(\exists j: w^*_\alpha=w_j\).

### 2020-06-13 2020/OPTIMA/optima_paper.tex
check all paper

--- 2020-06-13 2020/OPTIMA/optima_paper.tex
@@
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Parallel global optimization algorithm with uniform convergence}
 % If the paper title is too long for the running head, you can set
@@
 
-The feasible domain \(G\) can also be defined by the constraints that essentially complicate the problem.
+The feasible domain can also be defined by the constraints that essentially complicate the problem.
 The problem statement in this case will take the following form:
@@
   \label{eq:constrained_problem}
-  \varphi(y^*)=\min\{\varphi(y):y\in G=D \cap \{y: g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}\}.
+  \varphi(y^*)=\min\{\varphi(y):y\in D \cap \{y: g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}\}.
 \end{equation}
@@
   \label{eq:many_problems}
-  \min\left\{\varphi_1(y), y\in G_1 \right\}, \min\left\{\varphi_2(y), y\in G_2\right\},...,
-\min\left\{\varphi_q(y), y\in G_q\right\}.
+  \min\left\{\varphi_1(y), y\in D_1 \right\}, \min\left\{\varphi_2(y), y\in D_2\right\},...,
+\min\left\{\varphi_q(y), y\in D_q\right\}.
 \end{equation}
@@
     \psi (x^{*})=\min_{x\in [0,1]}\psi (x), \\
-    \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu },&\nu <M\\(g_{M}(x)-g_{M}^{*})/H_{M},&\nu
-=M\end{cases}}
+    \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu },&\nu <M,\\(g_{M}(x)-g_{M}^{*})/H_{M},&\nu
+=M.\end{cases}}
   \end{array}
 \end{displaymath}
-
 Here \(M=\max_{}^{}\left\{\nu (x):x\in [0,1]\right\}\) and \(g_{M}^{*}=\min
@@
 has a solution. If \(M=m+1\), then \(g_{M}^{*}=\varphi(x^{*})\).
-The arcs of the functions \(\psi (x)\) are the H\"{o}lder ones within the sets
+The function \(\psi (x)\) satisfies the H\"{o}lder condition within the sets
 \(\cap _{i=0}^{j}Q_{i},0\leq j\leq M-1\)
-with the constant 1, and \(\psi (x)\) itself can have the breaks of the first kind at the boundaries
+with the constant 1, and \(\psi (x)\) can have jump discontinuities at the boundaries
 of these sets.
@@
 known in advance, these ones can be estimated in the course of solving the problem.
-
-A set of triples \(\{(x_k,\nu_k,z_k)\}, 1\leqslant k\leqslant n\) constitutes the search information
+A set of triples \(\{(x_k,\nu_k,z_k)\}, 1\leqslant k\leqslant n\), constitutes the search information
 accumulated by the method upon executing \(n\) steps.
@@
 includes the boundary points of the interval \([0,1]\) as well as the points of preceding trials by
-the lower indices in the order of increasing coordinate values i. e.
+the lower indices in the order of increasing coordinate values, i.e.
 \begin{equation}
   \label{eq:points}
-  0=x_0<x_1<\dotsc<x_{k+1}=1
+  0=x_0<x_1<\dotsc<x_{k+1}=1,
 \end{equation}
@@
 
-Step 2. For each integer number \(\nu ,1\leqslant \nu \leqslant m+1\) determine the corresponding
+Step 2. For each integer number \(\nu ,1\leqslant \nu \leqslant m+1\), determine the corresponding
 set \(I_{\nu }\) of the lower indices of the points, which the values of the functions \(g_{\nu
@@
 \end{displaymath}
-determine the maximum value of the index \(M=\max\{\nu (x_{i}),1\leq i\leq k\}\).
+and determine the maximum value of the index \(M=\max\{\nu (x_{i}),1\leq i\leq k\}\).
 
@@
   \label{step2}
-  \mu _{\nu }=\max\{\frac{|g_{\nu }(x_{i})-g_{\nu }(x_{j})|}{(x_{i}-
-x_{j})^{\frac{1}{N}}}:i,j\in I_{\nu },i>j\}.
+  \mu _{\nu }=\max \left\{ \frac{|g_{\nu }(x_{i})-g_{\nu }(x_{j})|}{(x_{i}-
+x_{j})^{\frac{1}{N}}}:i,j\in I_{\nu },i>j \right\}.
 \end{equation}
@@
   \label{eq:step_4}
-  z_{\nu }^{*}={\begin{cases}\min\{g_{\nu }(x_{i}):x_{i}\in I_{\nu }\}&\nu =M\\-\varepsilon
-_{\nu }&\nu <M\end{cases}},
+  z_{\nu }^{*}={\begin{cases}\min\{g_{\nu }(x_{i}):x_{i}\in I_{\nu }\},&\nu =M,\\-\varepsilon
+_{\nu },&\nu <M,\end{cases}}
 \end{equation}
@@
 
-Step 5. For each interval \((x_{i-1};x_{i}),1\leqslant i\leqslant k\) compute the characteristic
+Step 5. For each interval \((x_{i-1},x_{i}),1\leqslant i\leqslant k\) compute the characteristic
 \begin{equation}
@@
   R(i)={\begin{cases}\Delta _{i}+{\frac {(z_{i}-z_{i-1})^{2}}{(r_{\nu }\mu _{\nu
-})^{2}\Delta _{i}}}-2{\frac {z_{i}+z_{i-1}-2z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu
-(x_{i})=\nu (x_{i-1})\\2\Delta _{i}-4{\frac {z_{i-1}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu
-}}}&\nu =\nu (x_{i-1})>\nu (x_{i})\\2\Delta _{i}-4{\frac {z_{i}-z_{\nu }^{*}}{r_{\nu }\mu
-_{\nu }}}&\nu =\nu (x_{i})>\nu (x_{i-1})\end{cases}}
+})^{2}\Delta _{i}}}-2{\frac {z_{i}+z_{i-1}-2z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}},&\nu =\nu
+(x_{i})=\nu (x_{i-1}),\\2\Delta _{i}-4{\frac {z_{i-1}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu
+}}},&\nu =\nu (x_{i-1})>\nu (x_{i}),\\2\Delta _{i}-4{\frac {z_{i}-z_{\nu }^{*}}{r_{\nu }\mu
+_{\nu }}},&\nu =\nu (x_{i})>\nu (x_{i-1}),\end{cases}}
 \end{equation}
 where \(\Delta_{i}=(x_{i}-x_{i-1})^{\frac{1}{N}}\).
-The magnitudes \(r_{\nu }>1,\nu ={\overline {1,m}}\) are the parameters of the algorithm.
+The values \(r_{\nu }>1,\nu ={\overline {1,m}}\) are the parameters of the algorithm.
 The products \(r_{\nu}\mu_{\nu}\) used in computing the characteristics as the estimates of the
@@
 \label{step4}
-t=\argmax_{1\leqslant i \leqslant k+1}R(i)
+t=\arg \max_{1\leqslant i \leqslant k+1}R(i).
 \end{equation}
 
-Step 7. Perform the next trial in the middle of the interval \((x_{t-1};x_{t})\) if the indices of its
+Step 7. Perform the next trial in the middle of the interval \((x_{t-1},x_{t})\) if the indices of its
 end points are not the same: \(x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})\).
@@
 \(\varepsilon>0\) is a predefined precision.
-The value
+The values
 \begin{equation}
-\varphi_k^*=\min_{1\leqslant i \leqslant k}\varphi(x_i), x_k^*=\argmin_{1\leqslant i \leqslant
+\varphi_k^*=\min_{1\leqslant i \leqslant k}\varphi(x_i), \; x_k^*=\arg \min_{1\leqslant i \leqslant
 k}\varphi(x_i)
 \end{equation}
-is accepted as the estimate of the global optimal solution.
+are accepted as the estimates of the global solution.
 
@@
     \item The problem (\ref{eq:constrained_problem}) has a solution.
-    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\) are Lipschitzian on \(D\) with
+    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
 respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
-    \item Beginning with some step (i. e. if \(k\) from (\ref{eq:points}) is sufficiently large),
-    the values \(\mu_\nu\) from \ref{step2} satisfy the inequalities
+    \item Beginning with some step (i.e. if \(k\) from (\ref{eq:points}) is sufficiently large),
+    the values \(\mu_\nu\) from (\ref{step2}) satisfy the inequalities
     \begin{equation}
@@
   \end{enumerate}
-  Then any limit point \(y\) of the sequence \(\{y_k\} = \{y(x_k)\}\) generated
+  Then any limit point \(\overline{y}\) of the sequence \(\{y_k\} = \{y(x_k)\}\) generated
   by the index method while solving the problerm
@@
   \label{rem:r1}
-  From the relationship between the Holder and Lipschitz constants and from the condition (\ref{eq:conv_cond}) it follows
+  From the relationship between the H\"{o}lder and Lipschitz constants and from the condition (\ref{eq:conv_cond}) it follows
   that the parameters \(r_\nu\) from (\ref{step3_1}) should satisfy the condition
@@
 
-\begin{theorem} (On convergence condiotions) Let the conditions of the Theorem \ref{th:single_conv}
-  are satisfied for all of \(q\) problems from the set \ref{eq:many_problems} separately.
+\begin{theorem} (On convergence conditions) Let the conditions of the Theorem \ref{th:single_conv}
+  are satisfied for all of \(q\) problems from the set (\ref{eq:many_problems}) separately.
   Then solving all of the \(q\) problems by the algorithm from Section \ref{sec:method} will
@@
   \begin{equation}
-    R(t) = \max\{R_\varphi(i),R_\psi(j)\}
+    R(t) = \max\{R_\varphi(i),R_\psi(j)\}.
   \end{equation}
 
-  Let the algorithm is solving tho problems and the iterations conunter is \(l = k + s, l=0,1,2,\dots\).
+  Let the algorithm is solving two problems and the iterations counter is \(l = k + s, l=0,1,2,\dots\).
   Then, since for all of the problems Theorem \ref{th:single_conv} is true, at least one of the sequences
@@
     \label{eq:th1}
-    R_\varphi(i)\leqslant\Delta_i(1 + \frac{1}{r_\nu^2}) - \frac{2(z_i+z_{i-1}-2z^*_\nu)}{r_\nu\mu_{\varphi,\nu}}
+    R_\varphi(i)\leqslant\Delta_i(1 + \frac{1}{r_\nu^2}) - \frac{2(z_i+z_{i-1}-2z^*_\nu)}{r_\nu\mu_{\varphi,\nu}}.
   \end{equation}
@@
 Also, in order to demonstrate the properties of the method to be preserved at essentially
-different properties of the problems, a mixed class consisting of 50 problems with two-
-dimensional GKLS functions and 50 problems with \(F_{GR}\) ones was generated.
+different properties of the problems, a mixed class consisting of 50 problems with 
+two-dimensional GKLS functions and 50 problems with \(F_{GR}\) ones was generated.
 Examples of contour plots of the considered problems are presented in Fig. \ref{fig:isolines}.
-The feasible area is highlighted by color.
+The feasible domain is highlighted by color.
 
@@
 A test problem was considered to be solved if the optimization method executes the next trial
-\(y^k\) in the \(\delta\)-proximity of the global minimum \(y^*\) i. e.
-\(\left\|y^k-y^*\right\|\leqslant \delta = 0.01\left\|b-a\right\|\)
-Where \(a\) and \(b\) are the left and right boundaries of the hypercube from (\ref{eq:task}).
+\(y^k\) in the \(\delta\)-vicinity of the global minimizer \(y^*\), i.e.
+\(\left\|y^k-y^*\right\|\leqslant \delta = 0.01\left\|b-a\right\|\), 
+where \(a\) and \(b\) are the left and right boundaries of the hypercube from (\ref{eq:task}).
 If this relation is not fulfilled upon reaching the limit of number of iterations, the problem was
@@
 Was taken into account as well.
-The mean maximum distance (in terms of \(l_{\inf}\)-norm) form the current estimate of the
+The mean maximum distance (in terms of \(l_{\infty}\)-norm) form the current estimate of the
 optimum to its actual position computed on the set of problems (\ref{eq:many_problems}):
@@
 The results of solving the test problems by the sequential and parallel versions of the modified
-IAGS for solving the sets of problems are presented in Table \ref{tab:speedup}.
+IAGS for solving the sets of problems are presented in Table~\ref{tab:speedup}.
 For all two-dimensional classes of problems, the parameter \(r=4.7\).
@@
 
-Let us consider a test problem proposed in \cite{BinhKorn1999}:
+Let us consider a test problem from \cite{BinhKorn1999}:
 \begin{equation}
@@
       \right .
-      y_1\in [-1;2],y_2\in [-2;1]
+      , y_1\in [-1,2],y_2\in [-2,1],
       \\s.t.
@@
       \begin{array}{l}
-        g_1(y) = (y_1 - 5)^2 + y_2^2 - 25 \leqslant 0 \\
-        g_2(y) = -(y_1 - 8)^2 - (y_2 + 3)^2 + 7.7 \leqslant 0\\
+        g_1(y) = (y_1 - 5)^2 + y_2^2 - 25 \leqslant 0, \\
+        g_2(y) = -(y_1 - 8)^2 - (y_2 + 3)^2 + 7.7 \leqslant 0.\\
       \end{array}
@@
 \((\lambda_1,\lambda_2)\) so that
-\(\lambda_1^i=i h,\: \lambda_2^i=1-\lambda_1^i,\: h=10^{-2},i=\overline{1, 100}\).
+\(\lambda_1^i=i h,\: \lambda_2^i=1-\lambda_1^i,\: h=10^{-2},\: i=\overline{1, 100}\).
 
@@
 featuring the density of the points approximating the Pareto set was computed also.
-\begin{displaymath}
+\begin{equation*}
+\begin {array}{l}
   SP(S)=\sqrt{\frac{1}{|S|-1} \sum_{i=1}^{|S|} (\overline{d}-d_i)^2},
-  \:\overline{d}=mean\{d_i\},\:d_i=\min_{s_i,s_j\in S:s_i\ne s_j}||F(s_i)-F(s_j)||_1,\: F=(f_1,f_2)
-\end{displaymath}
+  \; \overline{d}=mean\{d_i\},
+	\\d_i=\min_{s_i,s_j\in S:s_i\ne s_j}||F(s_i)-F(s_j)||_1,\; F=(f_1,f_2).
+	\end{array}
+\end{equation*}
 In the case of separate solving of the problem \(SP_{single}=0.984\) whereas when solving of
@@
 
-In the course of present work, the support of the nonlinear constraints was implemented in the
+In the course of present work, the support of the non-convex constraints was implemented in the
 algorithm solving a set of the global optimization problems.
-Also, the essential conditions of convergence for the developed method were obtained.
+Also, the sufficient conditions of convergence for the developed method were obtained.
 The numerical experiments demonstrating the advantages of the considered approach over
@@
 by reducing the costs of supporting the search information for the set of problems and, this way,
-improving the indices of speedup in time due to the parallelization.
+improving the indices of time speedup due to the parallelization.
 Also, it is planned to implement a version of considered algorithm working in the distributed

### 2020-06-13 2020/OPTIMA/optima_paper.tex
check all paper

@@
 \begin{proof}
-  Let us consider two random problems from the set \ref{eq:many_problems}.
+  Let us consider two random problems from the set (\ref{eq:many_problems}).
   \begin{equation}

### 2020-06-13 2020/OPTIMA/optima_paper.tex
OPTIMA: remove redundant paragraph

@@
 method among the problems dynamically.
-
-The theory of convergence of such an approach for the case of solving the problems without
-constraints is presented in \cite{BarkalovStrongin2018}.
-If there are constraints, the characteristics of the intervals, at the boundaries of which different
-numbers of constraints are violated are computed according to the bottom rows of
-(\ref{step3_1}).
-One can notice that even in this case the magnitudes of characteristics are normalized as well.
-In the case of exact estimates \(z_{\nu }^{*}\) and \(\mu _{\nu }\), the scale of these ones
-doesn't depend on the objective function and constraints of particular problem.
-So far, the speculations from \cite{BarkalovStrongin2018} can be conducted in the case of the
-use of the index scheme as well.
 

### 2020-06-14 2020/OPTIMA/optima_paper.tex
OPTIMA: fix issues with eps-reserves

@@
   \label{eq:step_4}
-  z_{\nu }^{*}={\begin{cases}\min\{g_{\nu }(x_{i}):x_{i}\in I_{\nu }\},&\nu =M,\\-\varepsilon
-_{\nu },&\nu <M,\end{cases}}
+  z_{\nu }^{*}={\begin{cases}\min\{g_{\nu }(x_{i}):x_{i}\in I_{\nu }\},&\nu =M,\\
+  0,&\nu <M,\end{cases}}.
 \end{equation}
-where the vector with the non-negative coordinates \(\varepsilon _{R}=(\varepsilon
-_{1},..,\varepsilon _{m})\) is called the vector of reserves.
 
@@
 GKLS \cite{Gaviano2003} allows obtaining the functions of predefined dimensionality and
-with predefined number of extrema.
+with predefined number of global optimums.
 In combination with GCGen, two sets of 100 problems each of the dimensionalities 2 and 3
@@
 For all two-dimensional classes of problems, the parameter \(r=4.7\).
-In the case of the three-dimensional problems, \(r=4.7,\: \varepsilon_\nu=0.1\).
+In the case of the three-dimensional problems, \(r=4.7\) and also to improve the convergence speed
+\(\varepsilon\)-reservation technique from \cite{Strongin2000} Chapter 8.3 with
+\(\varepsilon=0.1\) has been employed.
 In all experiments, an additional computational load was introduced into the objective function

### 2020-06-15 2020/OPTIMA/optima_paper.tex
minor corrections

@@
 \label{sec-intro}
-Global optimization problems with non-convex constraints are considered traditionally to be one
+Global optimization problems with non-convex constraints are considered to be one
 of the most difficult optimization problems.
- The finding of the global minimum of a function of several variables appears to be more
+Finding of the global minimum of a function of several variables appears to be more
 difficult often than a local optimization in a thousand-dimensional space.
@@
 \end{equation}
- is a classical scheme of reduction of the
-dimensionality of the initial problem for the global optimization algorithms
+ is the well-known scheme of dimensionality reduction
+ of the initial problem for the global optimization algorithms
 \cite{Sergeyev2013}.
@@
 \(x\in Q_{j},0\leqslant j<\nu ,x\notin Q_{\nu }\).
-The finding of the first violation of any constraint terminates the trial at the point \(x\).
+The first observed violation of any constraint terminates the trial at the point \(x\).
 In the case when the point \(x\) is feasible, i.e. \(x\in D\), the trial includes the computing of all
 functions of the problem.
-At that, the value of the index is accepted to be equal to the magnitude \(\nu =m+1\).
+At that, the value of the index is accepted as \(\nu =m+1\).
 The pair \(\nu =\nu (x), z=g_{\nu }(x)\) where the index \(\nu\) falls into the borders \(1\leqslant
@@
 constraints to an unconstrained problem of minimization of a discontinuous function:
-
 \begin{displaymath}
@@
   z_{\nu }^{*}={\begin{cases}\min\{g_{\nu }(x_{i}):x_{i}\in I_{\nu }\},&\nu =M,\\
-  0,&\nu <M,\end{cases}}.
+  0,&\nu <M.\end{cases}}
 \end{equation}
@@
   x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})-\operatorname {sgn}(z_{t}-z_{t-1}){\frac {|z_{t}-
-z_{t-1}|^{n}}{2r_{\nu }\mu _{\nu }^{n}}},\nu =\nu (x_{t})=\nu (x_{t-1}),
+z_{t-1}|^{N}}{2r_{\nu }\mu _{\nu }^{N}}},\nu =\nu (x_{t})=\nu (x_{t-1}),
 \end{displaymath}
@@
 \end{equation}
-  where \(y^*\) from (\ref{eq:constrained_problem}).
+  where \(y^*\) is from (\ref{eq:constrained_problem}).
 \end{theorem}
@@
 \begin{proof}
-  Let us consider two random problems from the set (\ref{eq:many_problems}).
+  Let us consider two random problems from the set (\ref{eq:many_problems})
   \begin{equation}
       \begin{array}{lr}
-        \min\{\varphi(y):y\in D_1,\: g_j^\varphi(y)\leqslant 0, 1\leqslant j\leqslant m_1\} \\
-        \min\{\psi(y):y\in D_2,\: g_j^\psi(y)\leqslant 0, 1\leqslant j\leqslant m_2\}
+        \min\{\varphi(y):y\in D_1,\: g_j^\varphi(y)\leqslant 0, 1\leqslant j\leqslant m_1\}, \\
+        \min\{\psi(y):y\in D_2,\: g_j^\psi(y)\leqslant 0, 1\leqslant j\leqslant m_2\}.
       \end{array}
@@
   \begin{displaymath}
-    R_\varphi(i)=2\Delta_i - 4\frac{z_i-z^*_\nu}{r_\nu\mu_{\varphi,\nu}}
+    R_\varphi(i)=2\Delta_i - 4\frac{z_i-z^*_\nu}{r_\nu\mu_{\varphi,\nu}}.
   \end{displaymath}
@@
       \Delta_j - \frac{2\Delta_j(w_{j-1}-w_j)}{r_\alpha\mu_{\psi,\alpha}\Delta_j} \geqslant
-      \Delta_j - \frac{\Delta_j}{r_\alpha} = \Delta_j\left(1-\frac{2}{r_\alpha}\right).
+      \Delta_j - \frac{\Delta_j}{r_\alpha} = \Delta_j\left(1-\frac{1}{r_\alpha}\right).
     \label{eq:th4}

### 2020-06-15 2020/OPTIMA/optima_paper.tex
OPTIMA: Fixes in formula and bib file

@@
 of several problems arises if the global optimization problem has a discrete parameter or, for
-example, in solving a multicriteria optimization problem using criteria convolution. The
+example, in solving a multicriteria optimization problem by applying scalarization techniques. The
 considered algorithm employs Peano-type space-filling curves for the reduction of the
@@
   If \(\nu(w_{j-1})=\nu(w_{j})=\alpha\) then the first branch of the rule (\ref{step3_1}) is applied and
-  \begin{dmath}
+  \begin{equation}
+    \begin{array}{l}
       R_\psi(j)=\Delta_j + \frac{(w_j-w_{j-1})^2}{r_\alpha^2\mu_{\psi,\alpha}^2\Delta_j}
-      - \frac{2(w_j+w_{j-1}-2w^*_\alpha)}{r_\alpha\mu_{\psi,\alpha}} \geqslant
-      \Delta_j - \frac{2(w_j+w_{j-1}-2w^*_\alpha)}{r_\alpha\mu_{\psi,\alpha}} =
-      \Delta_j - \frac{2\Delta_j(w_{j-1}-w_j)}{r_\alpha\mu_{\psi,\alpha}\Delta_j} \geqslant
-      \Delta_j - \frac{\Delta_j}{r_\alpha} = \Delta_j\left(1-\frac{1}{r_\alpha}\right).
+        - \frac{2(w_j+w_{j-1}-2w^*_\alpha)}{r_\alpha\mu_{\psi,\alpha}} \geqslant \\
+        \geqslant\Delta_j - \frac{2(w_j+w_{j-1}-2w^*_\alpha)}{r_\alpha\mu_{\psi,\alpha}} =
+        \Delta_j - \frac{2\Delta_j(w_{j-1}-w_j)}{r_\alpha\mu_{\psi,\alpha}\Delta_j} \geqslant\\
+        \geqslant \Delta_j - \frac{2\Delta_j}{r_\alpha} = \Delta_j\left(1-\frac{2}{r_\alpha}\right).
+      \end{array}
     \label{eq:th4}
-  \end{dmath}
+  \end{equation}
 

### 2020-06-20 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
OptLet sec 1-2, ref.

@@
 \usepackage{amsmath}
-\usepackage{hyperref}
+%\usepackage{hyperref}
 %\usepackage{academicons}
@@
 
-\title{A Computationally Efficient Approach for Solving Lexicographic Multicriteria Optimization Problems\thanks{This research was supported by the Russian Science Foundation, project No 16-11-10150.
+\title{Computationally Efficient Approach for Solving Lexicographic Multicriteria Optimization Problems\thanks{This research was supported by the Russian Science Foundation, project No 16-11-10150.
 %``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''.
@@
 \begin{abstract}
-This paper proposes a computationally efficient approach for solving complex lexicographic multicriteria optimization problems in which efficiency criteria can be multiextremal and computing the values of criteria can be time-consuming.
-It is also assumed that in the course of calculations, how problems are formulated may change and consequently, solving dynamically defined sets of multicriteria optimization problems may become necessary.
+In this paper, we propose a computationally efficient approach for solving complex multicriteria lexicographic optimization problems, which can be complicated by the multiextremal nature of the efficiency criteria and extensive volume of computations required to calculate the criteria values.
+The formulation of problems is assumed to be the subject to change as well, which, in turn, may require solving dynamically defined sets of multicriteria optimization problems.
 The proposed approach is based on reducing multidimensional problems to one-dimensional global optimization problems, utilizing efficient global search algorithms, and reusing the search information obtained in the process of calculations.
-Results of computational experiments demonstrate that the proposed approach can significantly reduce the computational complexity of solving multicriteria optimization problems.
-\keywords{Multicriteria optimization \and Lexicographic ordering \and Global optimization \and Dimensionality reduction \and Search information \and Computational complexity}
+The results of numerical experiments confirm that the proposed approach demonstrates high computational efficiency of solving multicriteria global optimization problems.
+\keywords{Multicriteria optimization \and Lexicographic ordering \and Global optimization \and Dimensional reduction \and Search information \and Computational complexity}
 \end{abstract}
@@
 
-Multicriteria optimization (MCO) problems are among the most common statements used to address decision-making problems and are widely found in applications.
-Multicriteria optimization is an area of vigorous scientific research, and during this time a large number of efficient methods of multicriteria optimization have been proposed and many practical problems have been solved (see, e.g., monographs \cite{c1,c2,c3,c4}) and reviews of scientific and practical results in this field \cite{c6,c7,c8,c9}.
+Multicriteria optimization (MCO) problems are among the most general statements used to address decision-making problems.
+These statements are commonly used in various applications.
+Multicriteria optimization is a dynamic area of scientific research.
+So far, researchers proposed a large number of efficient methods of multicriteria optimization and found solutions to many practical problems (see, e.g., monographs \cite{c1,c2,c3,c4,c5} and reviews \cite{c6,c7,c8,c9}).
 
-A key feature of MCO problems is the absence, in most cases, of a single decision which would best fulfill all efficiency criteria, due to possible inconsistencies.
-As a result, solving MCO problems usually requires finding several compromise (efficient and non-dominated) decisions that cannot be improved without damaging the effectiveness of some efficiency criteria in the MCO problem.
+A key aspect of MCO problems is the absence, in most cases, of a single decision, which would best satisfy all the efficiency criteria, since these criteria can be contradictory.
+As a result, solving MCO problems usually requires finding several compromise decisions (efficient and non-dominated) that cannot be improved without worsening the effectiveness of other efficiency criteria in the MCO problem.
 
-Among the approaches which have been developed for solving MCO problems, \textit{lexicographic optimization} methods can be singled out, when some ordering of criteria by importance is implemented and optimization is carried out based on their ordering \cite{c3}.
-\textit{Iterative methods} represent another approach \cite{c6,c10}, when the decision maker is actively involved in the process of selecting decisions.
-An actively appropriated approach consists in developing and applying evolutionary algorithms based on the simulation of certain natural phenomena to solve MCO problems \cite{c10,c11,c12,c13}.
-Among the widely used endeavors in solving MCO problems is the \textit{scalarization} approach, which applies methods involving the convolution of criteria to a single scalar criterion (see, e.g.
+Finding the whole set of efficient decisions may require a lot of computations and is often deemed redundant for the decision-maker.
+As a result, approaches that allow obtaining only particular limited sets of effective decisions are used in practice more often.
+
+Among these approaches is the scalarization approach, which is based on methods involving the convolution of criteria to a single scalar criterion (see, e.g.
 \cite{c2,c14}).
+Another approach relies on the iterative methods \cite{c6,c10} -- the decision-maker is actively involved in the process of selecting decisions.
+Another popular approach is to develop and apply evolutionary algorithms based on the simulation of certain natural phenomena to solve MCO problems \cite{c10,c11,c12,c13}.
+Another promising method is the application of new axiomatic of numerical infinities for the solution of the MCO problems \cite{c57,c58,c59}.
 
-This work is devoted to solving lexicographic multicriteria optimization (MCOlex) problems that arise when designing complex technical objects and systems.
-In such applications, efficiency criteria may have a complex \textit{multiextremal} form, and calculating criteria values may require \textit{large amounts of computations}.
-In such conditions, finding even one efficient decision requires significant calculations, while determining several (or a whole set) of efficient decisions becomes an issue of \textit{great computational complexity}.
+A note worthy approach to solving MCO problems is lexicographic optimization methods.
+Such methods involve ranking the criteria by the degree of their importance and using the obtained ranking for subsequent optimization \cite{c3}.
+When criteria are ranked by importance, an MCO problem can be solved via consecutive optimization of criteria in the order of their importance: the most important criterion is optimized first, followed by the optimization of the second criterion based on the analysis of the set of decisions, in which the first criterion reaches the best value.
+This sequence is simple and clear and allows using many effective scalar optimization algorithms.
+Such a multistage sequence can be reduced to solving one scalar optimization problem by linear convolution of criteria \cite{c38}.
+At the same time, this approach excessively reduces the number of calculated efficient decisions, which limits the number of options for the decision-maker.
+The $\varepsilon$-constraint scalarization approach can be used to expand the set of effective decisions that can be considered by the decision-maker.
+In this case, only the most important criterion is used for optimization, while the maximum permissible value is set for other criteria (i.e., all other criteria are transformed into MCO problem constraints) \cite{c39,c40}.
+Some researchers \cite{c41} describe the extended version of the $\varepsilon$-constraint method, in which violation of restrictions is allowed if it leads to a noticeable improvement of values by certain criteria of an MCO problem.
+Application of this approach is rather complex because the maximum allowable values (the concessions) of criteria must be set initially before calculations begin, which often causes substantial difficulties.
+A proposed solution \cite{c42} of this issue is to combine the initial solution sequence for lexicographical optimization problems, when concessions can be set in the process of calculations after each subsequent optimization stage (although such a scheme again requires the solution of the series of scalar optimization problems).
+The approach proposed in \cite{c43} allows even greater freedom in formulating problems of searching for effective decisions, when the order of criteria importance can also change dynamically.
+It is necessary to note that the presence of variability in problem statement for the search of effective decisions is very important as requirements to optimality of selected decisions can be varied by the decision-maker in the course of calculations.
+Such variability is also provided in other approaches to solving MCO problems by various methods of criteria scalarization -- see, for example, \cite{c31,c44}.
+In \cite{c1} a combined approach is proposed to the application of scalarization criteria and additional constraints of the $\varepsilon$-constraint method.
 
-The properties listed above highlight a key feature of the MCOlex problems   high computational complexity.
-One promising attempt to solve such problems is using the model-based approach, when, after a small number of calculations determining the values of time-consuming criteria and, fast-computable approximating functions are constructed \cite{c15,c16}.
-This approach is quite efficient; however, it is difficult to construct good approximations given the multiextremal behavior of efficiency criteria.
+This work focuses on solving lexicographic multicriteria optimization (MCOlex) problems that arise when designing complex technical objects and systems.
+In such applications, efficiency criteria may have a complicated multiextremal form; and calculating criteria values may require a large amount of computations.
+In such conditions, finding even one efficient decision requires a significant volume of calculations, while determining several  of efficient decisions becomes an issue of great computational complexity.
 
-This paper proposes a solution to the time-consuming class of MCOlex problems by using an approach based on the following main points.
-First of all, the problem is reduced to solving a sequence of global optimization problems with non-convex constraints (GCO) \cite{c2,c14}.
-Then, efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization are applied for solving the GCO problems \cite{c17,c18}.
-And finally, the search information obtained in the process of solving the MCOlex problem is fully utilized.
-In general, the developed approach allows us to significantly reduce the amount of executed computations, up to just a few iterations when searching for new efficient decisions.
+The properties listed above highlight a key feature of the MCOlex problems, namely, high computational complexity.
+One promising attempt to solve such problems is using the model-based approach, when calculation of a small number of time-consuming criteria values is followed by constructing the fast-computable approximating functions \cite{c15,c16}.This approach is quite efficient; however, it is difficult to construct adequate approximations given the multiextremal behavior of efficiency criteria.
+
+In this paper, we propose a solution to the time-consuming class of MCOlex problems based on the following main points.
+First, the problem is reduced to solving a sequence of global optimization problems with nonlinear constraints (GCO) \cite{c2,c14}.
+Then, the efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization are applied for solving the GCO problems \cite{c17,c18}.
+Finally, all the search information obtained in the process of solving the MCOlex problem is intensively reused.
+In general, the developed approach leads to a significant reduction of the amount of executed computations -- up to just a few iterations -- when searching for new efficient decisions.
 
 The further structure of the paper is as follows.
-Section \ref{sec:2} presents a new class of optimization problems -- multistage lexicographic multicriteria optimization (MMCOlex) problems -- whose solution is reduced to solving a sequence of global optimization problems with non-convex constraints.
-Section \ref{sec:3} presents the basics of the approach that has been developed: reducing multidimensional MCOlex problems to one-dimensional global optimization problems, applying efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization, and reusing the search information obtained in the process of calculations.
-Section \ref{sec:4} contains the results of numerical experiments.
-The paper concludes by discussing the results that were obtained and suggests some main directions for further research.
+Section \ref{sec:2} presents a new class of optimization problems–multistage lexicographic multicriteria optimization (MMCOlex) problems, the solution of which is reduced to solving a sequence of global optimization problems with nonlinear constraints.
+Section \ref{sec:3} considers the basics of the developed approach: reducing multidimensional MCOlex problems to one-dimensional global optimization problems, applying efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization, and reusing the search information obtained in the process of calculations.
+Section \ref{sec:4} provides a theoretical analysis of the effectiveness of reusing the search information for solving MMCOlex problems.
+Section \ref{sec:5} contains the results of numerical experiments.
+Discussion of the results and suggestions for further research conclude the paper.
 
@@
 
-The multicriteria (or vector) optimization (MCO) problem can be defined as follows:
+The multicriteria (or vector) optimization (MCO) problem can be formulated as:
 \begin{equation}\label{eq:1}
@@
 \end{equation}
-where $f(y) = (f_1(y), f_2(y), \dots , f_s(y))$ is the vector efficiency criterion, $y = (y_1, y_2, \dots , y_N)$ is the vector of variable parameters, and $N$ is the dimension of the multicriteria optimization problem to be solved.
-The search domain $D$ defines the set of possible parameter values and is usually an $N$-dimensional hypercube
+where $f(y) = (f1(y),f2(y),\dots,fs(y))$  is the vector efficiency criterion, $y = (y_1,y_2,\dots,y_N)$ is the vector of variable parameters, and $N$ is the dimension of the multicriteria optimization problem to be solved.
+The search domain $D$ defines the set of possible parameter values and commonly has a shape of an $N$-dimensional hypercube
 \begin{equation}\label{eq:2}
-D  = \{ y\in R^N: a_i \leq y_i \leq b_i, \; 1 \leq i \leq N \}
+D  = \{ y\in R^N: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}
 \end{equation}
@@
 
-Without a loss of generality, it is assumed that the values of the efficiency criteria are not negative and their decrease corresponds to an increase in the effectiveness of the considered decisions $y \in D$.
-Problem (\ref{eq:1}) is considered in relation to the most difficult decision-making problems in which the efficiency criteria $f_i(y)$, $1 \leq i \leq s$, can be significantly multiextremal, and the procedures for calculating the values of efficiency criteria at points of the search domain $y \in D$ can be computationally expensive.
-It is also assumed that the criteria $f_i(y)$, $1 \leq i \leq s$, satisfy the Lipschitz condition
-\begin{equation}\label{eq:2l}
-|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, \; y',y''\in D,\; 1 \leq i \leq s,
+Without loss of generality, the values of the efficiency criteria are assumed to be non-negative, and their decrease corresponds to an increase in the effectiveness of the considered decisions $y \in D$.
+Problem (\ref{eq:1}) is considered in the context of the most complex decision-making problems, in which the efficiency criteria $f_i(y)$, $1 \leq i \leq s$, can be significantly multiextremal, and the procedures for calculating the values of the efficiency criteria at points of the search domain $y \in D$ can be computationally expensive.
+The criteria $f_i (y)$, $1 \leq i \leq s$ are also assumed to satisfy the Lipschitz condition
+\begin{equation}\label{eq:3}
+|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, y',y''\in D, 1 \leq i \leq s,
 \end{equation}
-where $L_i$ is the Lipschitz constants for the function $f_i(y)$, $1 \leq i \leq s$, and $\|*\|$ denotes the Euclidean norm in $R^N$.
-Fulfillment of the Lipschitz condition means that for small variations of the decision $y \in D$, corresponding changes in the values of the criteria $f_i(y)$, $1 \leq i \leq s$, are limited.
+where $L_i$ is the Lipschitz constant for the function $f_i(y)$, $1 \leq i \leq s$, and ${\|*\|}$ denotes the Euclidean norm in $R^N$.
+The Lipschitz condition can be considered satisfied if for any variations of the decision $y \in D$ corresponding changes in the values of the criteria $f_i(y)$, $1 \leq i \leq s$, are limited.
 
-In problem (\ref{eq:1}), the efficiency criteria are usually contradictory and there is no decision $y \in D$, that would provide the best (smallest) values for all criteria at the same time, i.e.
-\begin{equation}\label{eq:3}
+In the problem (\ref{eq:1}), the efficiency criteria are usually contradictory and there is no decision $y \in D$ that would provide the best (smallest) values for all criteria at the same time, i.e.,
+\begin{equation}\label{eq:4}
 \nexists y^*\in D: f_i(y^*) = \min_{y \in D} {f_i (y)} , 1 \leq i \leq s.
 \end{equation}
-
-If the relation (\ref{eq:3}) is valid when solving the MCO problem, then decisions $\widetilde{y}^* \in D$ can be determined, at which the values of the criteria cannot be improved without deterioration of the values of some criteria $f_i(y)$, $1 \leq i \leq s$, that is
-\begin{equation}\label{eq:4}
-\nexists y'\in D: (f_i(y') \leq f_i(\widetilde{y}^*)), 1 \leq i \leq s, (\exists j : f_j(y') < f_j(\widetilde{y}^*)).
+If the relation (\ref{eq:4}) is valid in the solution of the MCO problem, then such decisions $\widetilde{y}^* \in D$ can be determined, at which the values of particular criteria cannot be improved without worsening the values of other criteria $f_i(y)$, $1 \leq i \leq s$,
+\begin{equation}\label{eq:5}
+\nexists y'\in D: (f_i(y') \leq f_i(\widetilde{y}^*), 1 \leq i \leq s), (\exists j: f_j(y') \leq f_j(\widetilde{y}^*)).
 \end{equation}
 
-Such non-improvable decisions are called Pareto \textit{efficient} or \textit{optimal}.
+Such unimprovable decisions are called Pareto efficient or optimal.
 
-The Pareto set $P(f,D)$ of all efficient decisions can be quite large, which makes it difficult for the decision maker to analyze.
-A possible way to reduce the number of efficient decisions under consideration is to assume that the effectiveness criteria are ordered in terms of importance, which is often the case in practical applications.
+The Pareto set $P(f,D)$ of all the efficient decisions can be quite large, which complicates its analysis by the decision-maker.
+A possible way to reduce the number of efficient decisions is to assume that the efficiency criteria are ordered by their importance, which is often the case in practical applications.
 
-Without a loss of generality, we will further assume that the criteria $f_i(y)$, $1 \leq i \leq s$, are ordered in importance according to their enumeration in (\ref{eq:1}).
-The ordering of the criteria determines the relative linear order in the search domain $D$, that is
-\begin{equation}\label{eq:4d}
+Without loss of generality, we will further assume that the criteria $f_i(y)$, $1 \leq i \leq s$, are ordered by their level of importance according to their enumeration in (\ref{eq:1}).
+Ranking of the criteria determines the relative linear order in the search domain $D$, that is
+\begin{equation}\label{eq:6}
 y' \prec y'' \Leftrightarrow \exists i, 1 \leq i \leq s: (f_i(y')<f_i(y'')) \wedge [ \forall j, 1 \leq j <i \Rightarrow (f_j(y')=f_j(y'')) ].
@@
 
-As a result, problem (\ref{eq:1}) is reduced to the lexicographic multicriteria optimization problem (MCOlex)
-\begin{equation}\label{eq:5}
+As a result, the problem (\ref{eq:1}) is reduced to the lexicographic multicriteria optimization problem (MCOlex)
+\begin{equation}\label{eq:7}
 f(y) = (f_1(y), f_2(y), \dots , f_s(y)) \to min_{lex},  y \in D,
 \end{equation}
-the solution of which is carried out in stages: initially, the first (most important) criterion is optimized; next, if the solution is not the only one, then the second criterion is optimized on the set of solutions of the first stage, and so on.
-\begin{equation}\label{eq:6}
-D_0=D,\; D_1=Arg \min_{y \in D_0}{f_1(y)},\dots, D_i=Arg \min_{y \in D_{i-1}}{f_i(y)},1 \leq i \leq s,
+which is then solved in stages: initially, the first (most important) criterion is optimized; next, if the solution is found to be non-unique, the second criterion is optimized on the set of solutions of the first stage, and so on.
+\begin{equation}\label{eq:8}
+D_1=Arg \min_{y \in D}{f_1(y)},\dots, D_i=Arg \min_{y \in D_{i-1}}{f_i(y)},1 \leq i \leq s,D_0=D,
 \end{equation}
-where \textit{Arg} means the set of all decisions that achieve the minimum value of the optimized criterion.
-It should be noted that the sequence of steps in (\ref{eq:6}) may not be fully completed if at some stage $l$, $1 \leq l \leq s$, the set $D_l$ degenerates and contains only a single decision $y \in D$.
+where $Arg$ is the set of all decisions that achieve the minimum value of the optimized criterion.
+It should be noted that the sequence of steps in (\ref{eq:8}) may not be fully completed if at a certain stage $l$, $1 \leq l \leq s$, the set $D_l$ degenerates and contains only a single decision $y \in D$.
 
-The set of efficient decisions $P_{lex}(f,D)$ obtained as a result of (\ref{eq:6}) is a subset of the Pareto domain $P(f,D)$ and can contain from one to several decisions $\widetilde{y}^* \in D$ in the case that the minimum value of the last criterion $f_s(y)$ is achieved at several points in the search domain $D$.
-Such a sharp reduction in the set of the efficient decisions being considered may be undesirable.
-Expanding the set of $P_{lex}(f,D)$ decisions can be achieved through somewhat weakening the ``strict'' lexicographic order and by allowing, at each stage of the computational scheme (\ref{eq:6}), the choice of decisions in a certain neighborhood of the minimum values of the optimized criterion, that is
-\begin{equation}\label{eq:7}
+The set of efficient decisions $P_{lex}(f,D)$ obtained as a result of (\ref{eq:6}) is a subset of the Pareto set $P(f,D)$ and can contain from one to several decisions $\widetilde{y}^* \in D$ in the case that the minimum value of the last criterion $f_s(y)$ is achieved at several points in the search domain $D$.
+Such a sharp reduction in the set of efficient decisions can be undesirable.
+Expanding the set of $P_{lex}(f,D)$ decisions can be achieved by somewhat of a weakening of the ``strict'' lexicographic order and by allowing the decisions to be selected in a certain neighborhood of the minimum values of the optimized criterion at each stage of the computational scheme (\ref{eq:8})
+
+\begin{equation}\label{eq:9}
 \begin{matrix}
@@
 
-This approach is also widely known as the method of successive concessions (MSC) \cite{c3,c4,c5}.
-The choice of feasible concessions $\delta_i$, $1 \leq i < s$, in the computational scheme (\ref{eq:7}) allows one to obtain any efficient decision $\widetilde{y}^* \in P_{lex}(f,D)$ and take into account the peculiarities of the MCOlex problem being solved.
-At the same time, this approach makes it necessary, at each stage of scheme (\ref{eq:7}), to solve more complex global optimization problems with non-convex constraints.
+This approach is also widely known as the $\varepsilon$-constraint method (or the method of successive concessions) \cite{c39,c40,c41,c42}.
+The choice of feasible concessions $\delta_i$, $1 \leq i \leq s$ in the computational scheme (\ref{eq:9}) allows obtaining any efficient decision $\widetilde{y}^* \in P_{lex}(f,D)$ and takes into account the peculiarities of the MCOlex problem.
+At the same time, this approach makes it necessary, at each stage of the scheme (\ref{eq:9}), to solve more complex global optimization problems with nonlinear constraints.
 
-It should also be noted that during the calculation process, it may be necessary to change the selected values of the feasible concessions $\delta_i$, $1 \leq i < s$; the concessions may be quite rigid or, conversely, excessively large.
-In general, one may need to change the order of importance in efficiency criteria.
-Such assumptions necessitate a more general formulation of how to solve the MCOlex problem and to provide possible solutions to many problems of the form (\ref{eq:7})
-\begin{equation}\label{eq:8}
+It should also be noted that during the calculation process, it may be necessary to change the selected values of the feasible concessions $\delta_i$, $1 \leq i \leq s$, the concessions may be quite rigid or, conversely, excessively large.
+In general, the change in the order of efficiency criteria importance may be required.
+Such assumptions necessitate a more general formulation of how to solve the MCOlex problem and to provide possible solutions to many problems of the form (\ref{eq:9})
+\begin{equation}\label{eq:10}
 \mathbb{P}_t=\{ P_1,P_2,\dots,P_t\},
@@
 which can be changed dynamically during calculations by adding new or deleting existing multicriteria optimization problems.
-In general, such an approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex).
+In general, this approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex).
 
@@
 Such heterogeneous coverage of the search domain $D$ can be constructed adaptively in the following manner: when choosing points for the next iterations, the global search is performed taking into account all the available search information (points of previous search iterations and values of the optimized function at these points), that is
-\begin{equation}\label{eq:9}
+\begin{equation}\label{eq:9q}
 y^{k+1}=G_k(\omega_k), \omega_k=\{(y^i,f^i=f(y^i)): 1 \leq i \leq k \},
@@
 As a result of this reduction, the initial multidimensional multicriteria optimization problem (\ref{eq:5}) is reduced to a one-dimensional problem:
-\begin{equation}\label{eq:10}
+\begin{equation}\label{eq:10q}
 f(y(x)) = (f_1(y(x)), f_2(y(x)), \dots , f_s(y(x)) \to min_{leq},  x \in [0,1].
@@
 \end{equation}
-where the constants $H_i$ are determined by the ratio $H_i=2L_i\sqrt{N+3}$, $1 \leq i \leq m$, $L_i$ are the Lipschitz constants from (\ref{eq:2l}), and $N$ is the dimension of the optimization problem (\ref{eq:1}).
+where the constants $H_i$ are determined by the ratio $H_i=2L_i\sqrt{N+3}$, $1 \leq i \leq m$, $L_i$ are the Lipschitz constants from (\ref{eq:3}), and $N$ is the dimension of the optimization problem (\ref{eq:1}).
 
@@
 
+\bibitem{c38} Zarepisheh, M., Khorram, E.: On the transformation of lexicographic nonlinear multiobjective  programs to single objective programs. Math. Methods Oper. Res. 74. pp. 217--231 (2011)
+\bibitem{c39} Haimes, Y.Y., Lasdon, L., Wismer, D.: On a bicriterion formulation of the problem of integrated systems identification and system optimization.
+IEEE Trans.
+Syst.
+Man Cybern.
+1.
+pp.
+296--297 (1971)
+\bibitem{c40} Chankong, V., Haimes, Y.Y.: Multiobjective decision making: theory and methodology. North-Holland Series in System Science and Engineering. New York (NY): Elsevier. (1983)
+\bibitem{c41} Ehrgott, M., Ruzika, S.: Improved $\varepsilon$-constraint method for multiobjective programming. J. Optim. Theory Appl. 138. pp. 375--396 (2008)
+\bibitem{c42} Rastegar, N., Khorram, E.: Relaxation of constraints in lexicographic multiobjective programming problems.
+A Journal of Mathematical Programming and Operations Research, 64.
+pp.
+2111--2129 (2015)
+\bibitem{c43} Castro-Gutierrez, J., Landa-Silva, D., Pérez, J.M.: Improved Dynamic Lexicographic Ordering for Multi-Objective Optimisation.
+In: Schaefer R., Cotta C., Kolodziej J., Rudolph G.
+(eds) Parallel Problem Solving from Nature, PPSN XI.
+PPSN 2010.
+Lecture Notes in Computer Science, 6239.
+pp 31-40 (2010)
+\bibitem{c44} Charles Audet, Gilles Savard, and Walid Zghal: Multiobjective Optimization Through a Series of Single-Objective Formulations SIAM J. Optim.,19(1). pp. 188--210 (2008)
+\bibitem{c45} Sergeyev, Y.D., Kvasov, D.E.: Deterministic Global Optimization: An Introduction to the Diagonal Approach (SpringerBriefs in Optimization). Springer. (2017)
+\bibitem{c46} Carr, C.R., Howe, C.W.: Quantitative Decision Procedures in Management and Economic: Deterministic Theory and Applications. New York: McGraw–Hill (1964)
+\bibitem{c47} Dam, E.R., Husslage, B., Hertog, D.: One-dimensional Nested Maximin Designs. Journal of Global Optimization, 46(2). pp. 287--306 (2010)
+\bibitem{c48} Gergel, V., Kozinov, E.: Efficient multicriterial optimization based on intensive reuse of search information, J. Glob. Optim., 71(1). pp. 73--90 (2018). doi:10.1007/s10898-018-0624-3.
+\bibitem{c49} Mart\'l, L., Garc\'la, J., Berlanga, A., Molina, J.M.: A stopping criterion for multi-objective optimization evolutionary algorithms, Information Sciences 367, pp. 700--718 (2016)
+\bibitem{c50} Dilettoso, E., Rizzo, S.A., Salerno, N.: A weakly pareto compliant quality indicator. Mathematical and Computational Applications 22(1), 25 (2017)
+\bibitem{c51} Wu, J., Azarm, S.: Metrics for quality assessment of a multiobjective design optimization solution set, Journal of Mechanical Design 123(1). pp. 18--25 (2001)
+\bibitem{c52} Audet, C., Bigeon, J., Cartier, D., Digabel,  S. Le, Salomon, L.: Performance indicators in multiobjective optimization. G-2018-90, (2018)
+\bibitem{c53} Huband, S., Hingston, P., Barone, L., and While, L.: A Review of Multiobjective Test Problems and a Scalable Test Problem Toolkit.
+IEEE Transactionson Evolutionary Computation, 10(5).
+pp.
+477--506 (2006)
+\bibitem{c54} Hansen, N., Auger, A., Finck, S., and Ros, R.: Real-Parameter Black-Box Optimization Benchmarking 2009: Experimental Setup.
+INRIA Research Report RR-6829, INRIA Saclay—Ile-de-France.
+updated February 2010 (2009)
+\bibitem{c55} Cust\'odio, A.L., Aguilar, J.F. Madeira, Ismael Vaz, Lu\'ls N. Vicente: Direct Multisearch for Multiobjective Optimization. IAM Journal on Optimization 21(3). pp. 1109-1140 (2011)
+\bibitem{c56} Dimo Brockhoff, Tea Tusar, Anne Auger, Nikolaus Hansen: Using Well-Understood Single-Objective Functions in Multiobjective Black-Box Optimization Test Suites.
+(2019) https://arxiv.org/abs/1604.00359
+\bibitem{c57} Cococcioni, M., Cudazzo, A., Pappalardo, M., Sergeyev, Y.D.: Solving the Lexicographic Multi-Objective Mixed-Integer Linear Programming Problem using branch-and-bound and grossone methodology, Communications in Nonlinear Science and Numerical Simulation, 84.
+pp.
+105177 (2020)
+\bibitem{c58} Cococcioni, M., Pappalardo, M., Sergeyev, Y.D.: Lexicographic multi-objective linear programming using grossone methodology: Theory and algorithm, Applied Mathematics and Computation, 318.
+pp.
+298--311 (2018)
+\bibitem{c59} Lai, L.
+, Fiaschi, L.
+, Cococcioni, M.: Solving Mixed Pareto-Lexicographic Multi-Objective Optimization Problems: The Case of Priority Chains, Swarm and Evolutionary Computation, 55.
+pp.
+100687 (2020) DOI: 10.1016/j.swevo.2020.100687
+
 \end{thebibliography}

### 2020-06-21 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
OptLet перевод в tex

@@
 
-\section{Problem statement}
+\section{Problem Statement}
 \label{sec:2}
@@
 \end{equation}
-where $f(y) = (f1(y),f2(y),\dots,fs(y))$  is the vector efficiency criterion, $y = (y_1,y_2,\dots,y_N)$ is the vector of variable parameters, and $N$ is the dimension of the multicriteria optimization problem to be solved.
+where $f(y) = (f_1(y),f_2(y),\dots,f_s(y))$  is the vector efficiency criterion, $y = (y_1,y_2,\dots,y_N)$ is the vector of variable parameters, and $N$ is the dimension of the multicriteria optimization problem to be solved.
 The search domain $D$ defines the set of possible parameter values and commonly has a shape of an $N$-dimensional hypercube
@@
 \begin{equation}\label{eq:5}
-\nexists y'\in D: (f_i(y') \leq f_i(\widetilde{y}^*), 1 \leq i \leq s), (\exists j: f_j(y') \leq f_j(\widetilde{y}^*)).
+\nexists y'\in D: (f_i(y') \leq f_i(\widetilde{y}^*), 1 \leq i \leq s), (\exists j: f_j(y') < f_j(\widetilde{y}^*)).
 \end{equation}
@@
 
-\section{The Approach: Key Ideas and Methods}
+\section{The Proposed Approach: Key Ideas and Methods}
 \label{sec:3}
 
-The extended formulation of lexicographic multicriteria optimization problems (\ref{eq:7})--(\ref{eq:8}) involves multiple solutions of multiextremal optimization problems with non-convex constraints.
-Problems of this type are computationally time-consuming and subject to the ``curse of dimensionality'': computational complexity increases exponentially with increasing dimensionality.
-The proposed approach for solving such problems is based on three key ideas:
+The extended formulation of lexicographic multicriteria optimization problems (\ref{eq:7})--(\ref{eq:8}) involves multiple solutions for global optimization (GO) problems with nonlinear constraints.
+Problems of this type are computationally time-consuming and subject to the ``curse of dimensionality'' -- computational complexity increases exponentially with increasing dimensionality.
+
+For many years, global optimization has been the subject of many scientific studies.
+The results of research have been published in a large number of monographs -- see, for example, \cite{c17,c18,c19,c20,c21,c22,c23,c24,c25}.
+Making a reasonable choice of the approach for solving GO problems from the entire set of already existing global search algorithms requires taking into account the following:
+
 \begin{itemize}
-  \item reduction of multidimensional problems of multiextremal optimization to problems of one-dimensional global search by using mappings based on Peano curves;
-  \item using the index method of global optimization, which allows solving optimization problems with constraints without using penalty functions;
-  \item repeated use of search information obtained in the process of computing, for solving multiple global optimization problems.
+	\item Since the estimate of the global minimum is an integral characteristic of the optimization problem (the value of the minimized function in the point of the global minimum should be compared with the values of the function in all other points of the search domain), the global search methods should provide at least some degree of coverage of the search domain,
+	\item Due to the ``curse of dimensionality'' reducing computational complexity requires the resulting coverage of the search domain to be heterogeneous (nonuniform), i.e.
+it must be dense only in the vicinity of globally optimal points,
+	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- on eof the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}),
+	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function in these points), which significantly increases the complexity and computational cost of the global optimization methods.
+Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47}, the Peano space-filling curves \cite{c17,c23},
+	\item When solving multiple problems of multistage lexicographic multicriteria optimization, global search methods can reduce the complexity of calculations by using search information obtained during the calculations.
 \end{itemize}
 
-\subsection{Dimension reduction of multidimensional multicriteria optimization problems}
+Given the requirements listed above, the proposed approach for solving MMCOlex problems is based on three key ideas:
 
-The search for numerical estimates of globally optimal points involves the construction of grids covering the search domain $D$ (see, e.g., \cite{c17,c18,c19,c20,c21,c22,c23,c24,c25}).
-This key feature of global optimization problems leads to the ``curse of dimensionality'', where the computational complexity of constructing coverage increases exponentially as the dimensions of optimization problems are enlarged.
-Significantly reduced complexity can be achieved by using heterogeneous grids that are denser only in the vicinity of globally optimal points.
-Such heterogeneous coverage of the search domain $D$ can be constructed adaptively in the following manner: when choosing points for the next iterations, the global search is performed taking into account all the available search information (points of previous search iterations and values of the optimized function at these points), that is
-\begin{equation}\label{eq:9q}
+\begin{itemize}
+	\item Reduction of multidimensional multiextremal optimization to problems of one-dimensional global search by the space-filling curves;
+	\item Using the global optimization methods that generate nonuniform coverages of the search domain;
+	\item Intensive search information obtained in the process of computations for solving multiple global optimization problems.
+\end{itemize}
+
+\subsection{Dimension Reduction of Multidimensional Multicriteria Optimization Problems}
+
+The nonuniform coverage of the search domain $D$ can be constructed adaptively in the following manner: when choosing points for the next optimization iterations, the global search is performed taking into account all the available search information (the points of previous search iterations and the values of the optimized function in these points), that is
+\begin{equation}\label{eq:11}
 y^{k+1}=G_k(\omega_k), \omega_k=\{(y^i,f^i=f(y^i)): 1 \leq i \leq k \},
 \end{equation}
-where $k$ is the number of the global search iterations being performed, $G_k$ is the decision rule of the applied global search algorithm, according to which the point $y^{k+1}$ of the next iteration is selected, and $\omega_k$ is the search information available at the $k$ step of the global search.
-
-Adaptive selection of points $y^{k+1} \in D$ at the iterations of the global search involves the analysis of volumetric multidimensional search information $\omega_k$, $k \gg 1$, which determines the high computational complexity of the decision rules $G_k$, $k=1,2,\dots$, of global search algorithms.
-This computational complexity can be significantly simplified by reducing the optimization problems to be solved using Peano \textit{space-filling curves} or \textit{evolvents} $y(x)$, that uniquely and continuously map the interval $[0,1]$ to an $N$-dimensional hypercube $D$ (see, e.g., \cite{c17,c18,c23}).
-As a result of this reduction, the initial multidimensional multicriteria optimization problem (\ref{eq:5}) is reduced to a one-dimensional problem:
-\begin{equation}\label{eq:10q}
-f(y(x)) = (f_1(y(x)), f_2(y(x)), \dots , f_s(y(x)) \to min_{leq},  x \in [0,1].
+where $k$ is the number of the global search iterations performed, $G_k$ is the decisive rule of the global search algorithm, according to which the point $y^{k+1}$ of the next iteration is selected, and $\omega_k$ is the search information available at the $k$ iteration of the global search.
+The computational complexity of the decisive rule can be significantly simplified by reducing the optimization problems by the Peano space-filling curves or evolvents $y(x)$ that uniquely and continuously map the interval [0,1] to an $N$-dimensional hypercube $D$ (see, e.g., \cite{c17,c18,c23}).
+As a result of this reduction, the initial multidimensional multicriteria optimization problem (\ref{eq:7}) is reduced to a one-dimensional problem:
+\begin{equation}\label{eq:12}
+f(y(x)) = (f_1(y(x)), f_2(y(x)), \dots, f_s(y(x))) \to min_{lex},  x \in [0,1].
 \end{equation}
@@
 It should be noted that the one-dimensional functions obtained as a result of the reduction satisfy the uniform H\"older condition (see \cite{c17,c18}), that is
-\begin{equation}\label{eq:11}
-|f_i (y(x'))-f_i (y(x''))| \leq H_i \|x'-x''\|, \; x',x''\in [0,1],\; 1 \leq i \leq s,
+\begin{equation}\label{eq:13}
+|f_i (y(x'))-f_i (y(x''))| \leq H_i |x'-x''|^{1/N}, x',x''\in [0,1], 1 \leq i \leq s,
 \end{equation}
-where the constants $H_i$ are determined by the ratio $H_i=2L_i\sqrt{N+3}$, $1 \leq i \leq m$, $L_i$ are the Lipschitz constants from (\ref{eq:3}), and $N$ is the dimension of the optimization problem (\ref{eq:1}).
+where the constants $H_i$ are determined by the ratio $H_i=2L_i\sqrt{N+3}$,  $1 \leq i \leq m$, the values $L_i$ are the Lipschitz constants from (\ref{eq:2}), and $N$ is the dimension of the optimization problem (\ref{eq:1}).
 
-Using dimension reduction the search information $\omega_k$ can be converted from (\ref{eq:9}) to the form
-\begin{equation}\label{eq:12}
+Using dimension reduction, the search information $\omega_k$ can be converted from (\ref{eq:11}) to 
+\begin{equation}\label{eq:14}
 A_k=\{(x_i, y(x_i), f_i=f(y(x_i))): 1 \leq i \leq k \},
 \end{equation}
-in which, in addition to the reduced representation, the data is arranged in ascending order\footnote{Data ordering is reflected by using a subscript.} of points $x_i$, $1 \leq i \leq k$, that is
-\begin{equation}
+where, in addition to the reduced representation, the data is arranged in ascending order\footnote{Data ordering is reflected by using a subscript.} of points $x_i$, $1 \leq i \leq k$, that is
+\begin{equation} \label{eq:15}
 x_1 < x_2 < \dots < x_k	
 \end{equation}
-for a more efficient execution of global search algorithms.
+for more efficient execution of global search algorithms.
 
-Dimension reduction using Peano curves is widely used in the development of efficient global search algorithms in the framework of the information-statistical theory of multiextremal optimization \cite{c17,c18,c23,c26,c27}.
-Numerical methods for constructing approximations of Peano curves (\textit{evolvents}) with a given accuracy are considered in \cite{c17,c18}.
+Dimension reduction by the Peano curves is widely used in the development of efficient global search algorithms in the framework of the information-statistical theory of multiextremal optimization \cite{c17,c18,c23,c26,c27}.
+Numerical methods for constructing approximations of the Peano curves (evolvents) with a given accuracy are considered in \cite{c17,c18}.
 
-\subsection{An efficient method for solving global optimization problems with 
-non-convex constraints}
+\subsection{An Efficient Method for Solving Global Optimization Problems with Nonlinear Constraints}
 
-The problems solved in the framework of the computational scheme (\ref{eq:7}) are one-dimensional global optimization problems with non-convex constraints
-\begin{equation}\label{eq:13}
-\min{\{\varphi(x):g_i(x)\leq 0, \; 1 \leq i \leq m,\; x\in \Delta=[0,1]\}}
+The problems solved in the framework of the computational scheme (\ref{eq:9}) are one-dimensional global optimization problems with nonlinear constraints
+\begin{equation}\label{eq:16}
+\min{\{\varphi(x):g_i(x)\leq 0, 1 \leq i \leq m, x\in \Delta=[0,1]\}}
 \end{equation}
-(the notation $g_{m+1}(x) = \varphi(x)$ will also be used later).
+(the expression $g_{m+1}(x) = \varphi(x)$ will also be used later).
 
-The presence of non-convex constraints significantly complicates the complexity of solving global optimization problems; the resulting solutions must belong to a feasible domain
-\begin{equation}\label{eq:14}
+The presence of nonlinear constraints significantly increasesthe complexity of global optimization problems; the resulting solutions must belong to a feasible domain
+\begin{equation}\label{eq:17}
 \Delta_g  = \{ x\in \Delta=[0,1]:g_i(x)\leq 0, 1 \leq i \leq m \}.
@@
 Various methods of approximating complex constraints with the aid of simpler forms (linear, convex, etc.) are widely used.
-However, the penalty function method is most commonly used, when calculating in an unfeasible region $\Delta \setminus \Delta_g$, some ``penalty'' is added to the values of the minimized function.
-But the penalty function method often involves repeatedly solving the optimization problem to determine a sufficient value of the penalty coefficient.
-More complete information on methods for solving optimization problems with constraints can be obtained, for example, in monographs \cite{c18,c26,c27,c28,c29}.
+However, the most commonly used method is the penalty function method -- when calculating in an unfeasible region $\Delta \setminus \Delta_g$, certain ``penalties'' are added to the values of the minimized function.
+However, the penalty function method often involves the repeated solution of the optimization problem to determine a sufficient value of the penalty coefficient.
+More complete information on methods for solving optimization problems with constraints is presented, for example, in monographs \cite{c28,c29}.
 
 The proposed approach is based on the original method of separate accounting for constraints developed in the framework of the information-statistical theory of multiextremal optimization \cite{c18}.
-The key idea of the approach consists in constructing a certain integral objective function without constraints, the solution of which also provides a solution to the initial constrained optimization problem (\ref{eq:13}); a more detailed presentation of this method is given below.
+The key idea of the approach is to construct an integral objective function without constraints, the solution of which also provides a solution to the initial constrained optimization problem (\ref{eq:16}); a more detailed presentation of this method is given below.
 
-We apply a sequential scheme to obtain the constraint values $g_i(y(x))$, $1 \leq i \leq m$, and the objective function $\varphi(x)$, according to which the constraint values are calculated strictly in the order of their numbers, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme).
-The number of constraints in which the values were calculated will be indicated below as the index $\nu=\nu(x)$, that is
-\begin{equation}\label{eq:15}
-\nu(x) : g_\nu(x)>0,\; g_j(x)\leq 0,\; 1 \leq j\leq \nu - 1,\; 1\leq \nu = \nu(x) \leq m
+To provide uniform calculations, a sequential scheme is applied to obtain values the constraints $g_i(y(x)) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$.
+According to this scheme, the constraint values are calculated strictly in their order, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme).
+The number of constraints, in which the values were calculated, will be referred below as the index $\nu=\nu(x)$, that is
+\begin{equation}\label{eq:18}
+\nu(x) : g_\nu(x)>0, g_j(x)\leq 0, 1 \leq j \leq \nu - 1, 1 \leq \nu = \nu(x) \leq m
 \end{equation}
- (when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index concept allows us to determine the integrated function for the problem (\ref{eq:13}) (see Fig. \ref{fig:1})
-\begin{equation}\label{eq:16}
+(when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index allows determining the integrated function for the problem (\ref{eq:16}) (see Fig. \ref{fig:1})
+\begin{equation}\label{eq:19}
 F(x)=g_\nu(x), \nu=\nu(x),
@@
 Its value at the point $x\in [0,1]$ is either the value of the left side of the constraint violated at this point (the case when $\nu \leq m$), or the value of the minimized function (the case when $\nu = m + 1$).
-The process of calculating the value of $F(x)$, $x\in [0,1]$, is completed either as a result of establishing the inequality $g_j(x)>0$, or as a result of reaching the value $\nu(x)=m+1$ (below, this procedure is called a \textit{trial}).
+The process of calculating the value of $F(x)$, $x\in [0,1]$, is completed either as a result of establishing the inequality $g_j(x)>0$, or as a result of reaching the value $\nu(x)=m+1$ (further in the paper, this procedure will be referred to as trial).
 
@@
   \includegraphics[width=\linewidth]{fig1}
-  \caption{An integrated function for the problem (\ref{eq:13}) using a sequential constraint calculation scheme: a) the function $F(x)$ from (\ref{eq:16}), b) the function $\Phi(x)$ from (\ref{eq:17})}
+  \caption{An example of the integrated function for the problem (\ref{eq:16}) using a sequential constraint calculation scheme: a) the function $F(x)$ from (\ref{eq:19}), b) the function $\Phi(x)$ from (\ref{eq:20})}
   \label{fig:1}
@@
 
-The function $F(x)$ consists of separate fragments of the constraints $g_i(x)$, $1 \leq i \leq m$, and the objective function $\varphi(x)$.
-For the homogeneity of the components, the function $F(x)$ is then converted to the form
-\begin{equation}\label{eq:17}
+The function $F(x)$ consists of separate fragments of the constraints ${g_i(x) \leq 0}$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$.
+For the homogeneity of the components, the function $F(x)$ is converted to the form
+\begin{equation}\label{eq:20}
 \Phi (x) = 
@@
 \end{equation}
-where $M$ is the constraint number for which there are no feasible points (the maximum value of the index is $\nu(x)$), $z^*_M$ is the value of the minimum constraint violation $g_M(x)$\footnote{If $M=m+1$, then $z_M^*$ is the minimum value of the function $\varphi(x)$.}, and $H_\nu$, $1\leq \nu \leq M$, are the H\"older constants from (\ref{eq:11}).
-As a rule, these values are a priori unknown, but when performing calculations, instead of these values, their adaptive estimates obtained in the process of calculations can be applied.
-For this, the search information $A_k$ from (\ref{eq:12}) should be expanded with data
-\begin{equation}\label{eq:18}
-A_k=\{(x_i, z_i, \nu_i, y(x_i), f_i=f( y(x_i))): 1 \leq i \leq k \},
+where $M$ is the constraint number, for which there are no feasible points (the maximum value of the index $\nu(x)$), $z^*_M$ is the value of the minimum constraint violation $g_M(x)$\footnote{If $M=m+1$, then $z_M^*$ is the minimum value of the function $\varphi(x)$.}, and $H_\nu$, $1\leq \nu \leq M$, are the H\"older constants from (\ref{eq:13}).
+As a rule, these values are a priori unknown, but when performing calculations, instead of these values, their adaptive estimates can be applied.
+For this, the search information $A_k$ from (\ref{eq:14}) should be expanded with data
+\begin{equation}\label{eq:21}
+A_k=\{(x_i, z_i, \nu_i, y(x_i), f_i=f(y(x_i))): 1 \leq i \leq k \},
 \end{equation}
-where $z_i= \Phi(x_i)$, $1 \leq i \leq k$. Then estimates of the required values can be determined as follows
-\begin{equation}\label{eq:19}
-M^k= \max{\{\nu=\nu(x_i), 1\leq i \leq k \}},
+where $z_i= \Phi(x_i)$, $1 \leq i \leq k$. Then the estimates of the required values can be determined as follows
+\begin{equation}\label{eq:22}
+M^k = \max{\{\nu=\nu(x_i), 1\leq i \leq k \}},
 \end{equation}
 
-\begin{equation}\label{eq:20}
+\begin{equation}\label{eq:23}
 z^k_\nu = 
  \begin{cases}
-   0, & \nu < M, \\
-   \min{\{g_\nu(x_i):\nu=\nu(x_i), 1\leq i \leq k \}}, & \nu = M,
+   0, & \nu < M^k, \\
+   \min{\{g_\nu(x_i):\nu=\nu(x_i), 1\leq i \leq k \}}, & \nu = M^k,
  \end{cases}
@@
 
-\begin{equation}\label{eq:21}
-h^k_l=\max{\{|z_i-z_j|/\sqrt[N]{x_i-x_j},\; \nu_i=\nu_j=l,\; i>j, \;1\leq l\leq M^k\}}.
+\begin{equation}\label{eq:24}
+h^k_l=\max{\{|z_i-z_j|/\sqrt[N]{x_i-x_j}, \nu_i=\nu_j=l, i>j, 1\leq l\leq M^k\}}.
 \end{equation}
 
-If $h_l^k$ in (\ref{eq:21}) turns out to be indefinite or equal to zero, then $h_l^k=1$ is taken.
-For the initial values of these quantities, $M^0 = H_l^0=1$, $z_l^0=0$, $1 \leq i \leq m+1$, can be taken.
+If $h_l^k$ in (\ref{eq:24}) turns out to be indefinite or equal to zero, then $h_l^k=1$ is taken.
+For the initial values of these parameters, $M^0 = H_l^0 = 1$, $z_l^0 = 0$, $1 \leq i \leq m+1$, can be taken.
 
-The general computational scheme of the \textit{global search algorithm} for multiextremal optimization problems with non-convex constraints \footnote{This method is also known as the index method.} (AGCS) is as follows \cite{c18}.
+In the framework of the proposed approach to minimize the function $\Phi(x)$, the global search algorithm for multiextremal optimization problems with nonlinear constraints\footnote{This method is also known as the index method -- see \cite{c18}.} (AGCS) is applied.
+The general computational scheme of this method can be presented as follows \cite{c18}.
 
-The first trial is carried out at an arbitrary internal point $x^1\in (0,1)$.
-The calculation results are constantly added to the set $A_k$ from (\ref{eq:12}); in addition, into $A_k$ are also added the points $x_0=0$ and $x_{k+1}=1$ (the values $z_0$ and $z_{k+1}$ are not defined and not used) for the convenience of subsequent formulations.
-The choice of the point $x^{k+1}$, $k\geq 1$, of any next trial is determined by the following rules.
+The first trial is carried out at an arbitrary internal point $x^1\in(0,1)$. The choice of the point $x^{k+1}$, $k \geq 1$ of any next trial is determined by the following rules
 
-\textit{Rule 1.} For each interval $(x_{i-1},x_i)$, $1 \leq i \leq k+1$, in the set $A_k$ calculate the characteristic $R(i)$, where
-\begin{equation}\label{eq:22}
-\begin{matrix}
-R(i) = 
- \begin{cases}
-   \rho_i + \frac{(z_i-z_{i-1})^2}{r^2_\nu h^2_\nu \rho_i} - 2 \frac{z_i+z_{i-1}-2z^*_\nu}{r_\nu h_\nu}, & \nu=\nu(x_{i-1})=\nu(x_i), \\
-   2 \rho_i - 4\frac{z_i-z^*_\nu}{r_\nu h_\nu}, & \nu=\nu(x_{i-1})<\nu(x_i), \\
-2 \rho_i - 4\frac{z_{i-1}-z^*_\nu}{r_\nu h_\nu}, & \nu=\nu(x_{i-1})>\nu(x_i), 
- \end{cases} \\
-\rho_j = \sqrt[N]{x_i-x_{i-1}}, 1\leq j\leq k+1.
-\end{matrix}
+\textit{Rule 1.} For each interval $(x_{i-1},x_i)$, $1 < i \leq k$, in the set $A_k$ calculate the characteristic
+
+\begin{equation}\label{eq:25}
+R(i)=\mathcal{R} (i,A_k),
 \end{equation}
-
-The constants $r_\nu>1$, $1\leq \nu \leq m+1$, are the parameters of the algorithm.
-The products $r_\nu h_\nu$ are used as an estimate of the H\"older constants $H_\nu$, $1\leq \nu \leq m+1$ from (\ref{eq:11}).
+where the expression $\mathcal{R}(i,A_k)$ is set by AGCS.
 
 \textit{Rule 2.} Determine the interval $(x_{t-1},x_t)$ that corresponds to the maximum characteristic
-\begin{equation}\label{eq:23}
-R(t) = \max{\{R(i):1 \leq i \leq k+1\}}.
+\begin{equation}\label{eq:26}
+R(t) = \max{\{R(i):1 < i \leq k\}}.
 \end{equation}
@@
 \textit{Rule 3.} Execute a new trial at the point $x^{k+1} \in (x_{t-1},x_t)$, determined in accordance with the expression
-\begin{equation}\label{eq:24}
-x^{k+1} = 
- \begin{cases}
-   \frac{x_t+x_{t-1}}{2}, & \nu(x_{t-1}) \neq \nu(x_t), \\
-   \frac{x_t+x_{t-1}}{2} + sign(z_t - z_{t-1}) \frac{1}{2r_\nu}[ \frac{|z_t - z_{t-1}|}{h_\nu}^N ], & \nu = \nu(x_{t-1}) = \nu(x_t).
- \end{cases}
+\begin{equation}\label{eq:27}
+x^{k+1} = \mathcal{X}(x_{t-1},x_t)
 \end{equation}
+where the expression $\mathcal{X}(x_{t-1},x_t)$ is also set by AGCS.
 
-Iterations of the algorithm are terminated if the stop condition is satisfied
-\begin{equation}\label{eq:25}
+Iterations of the algorithm are terminated if the stop criterion is satisfied
+
+\begin{equation}\label{eq:28}
 \rho_t\leq \varepsilon,
 \end{equation}
-where $t$ from (\ref{eq:23}) and $\varepsilon > 0$ is the given accuracy.
+where $t$ from (\ref{eq:26}), $\rho_t = \sqrt[N]{x_t-x_{t-1}}$ and $\varepsilon > 0$ is the given accuracy.
 
-Results of using the AGCS algorithm to solve the test problem from Fig.~\ref{fig:1} with the parameters $r_1=r_2=r_3=2$ and $\varepsilon=10^{-5}$ are shown in Fig.
-\ref{fig:2}.
-The coordinates of the trial points performed by the algorithm in the process of solving the problem are marked in Fig.
-\ref{fig:1} by three rows of vertical strokes.
-The strokes in the upper row correspond to points with a unit index, the second row to points whose indices equal to 2; the points marked with strokes in the bottom row belong to the feasible domain.
+Results of the application of the AGCS algorithm to solve the test problem from Fig.~\ref{fig:1} with the accuracy $\varepsilon=10^{-5}$ are shown in Fig.~\ref{fig:2}.
+The coordinates of the trial points performed by the algorithm in the process of solving the problem are marked in Fig.~\ref{fig:2} by three rows of vertical strokes.
+The strokes in the upper row correspond to points with a unit index, the second row to points, the indices of which equal to $\nu=2$; the points marked with strokes in the bottom row belong to the feasible domain.
 The coordinates of trials performed at close points are marked with a dark rectangle.
@@
 \begin{theorem}
-If the AGCS algorithm is used to solve problem (\ref{eq:13}), and the following conditions are satisfied:
+If the AGCS algorithm is used to solve the problem (\ref{eq:16}), and the following conditions are satisfied:
 \begin{enumerate}
-	\item the functions $g_j$, $1 \leq j \leq m+1$,  satisfy the H\"older condition of (\ref{eq:11}) with the constants $H_j$, $1 \leq j \leq m+1$,
-	\item for the quantities $h_\nu$ from (\ref{eq:21}), starting from a certain step, the inequalities are valid
-\begin{equation*}
-r_\nu h_\nu>2H_\nu, 1\leq \nu \leq m+1,
-\end{equation*}
+	\item the functions $g_j$, $1 \leq j \leq m+1$, $[0,1]$ satisfy the H\"older condition of (\ref{eq:13}) with the constants $H_j$, $1 \leq j \leq m+1$,
+	\item for the values $h_\nu$ from (\ref{eq:24}), starting from a certain step, the inequalities are valid\footnote{The values $r_{\nu} > 1$, $1 \leq \nu \leq m+1$,are the AGCS reliability parameters used for computing the interval characteristics in (\ref{eq:25}).}
+\begin{equation} \label{eq:29}
+r_\nu h_\nu > 2H_\nu, r_\nu > 1, 1 \leq \nu \leq m+1,
+\end{equation}
 \end{enumerate}
-then the set of limit points of the sequence $\{x^k\}$, generated by the algorithm coincides with the set of global minimum points of the problem (\ref{eq:13}) with $\varepsilon=0$ in the stop condition (\ref{eq:25}).
+then the set of limit points of the sequence $\{x^k\}$ generated by the algorithm coincides with the set of global minimum points of the problem (\ref{eq:16}) with $\varepsilon=0$ in the stopping criterion (\ref{eq:28}).
 \end{theorem}
 
-\subsection{Acceleration of computations based on the re-use of search information}
+\subsection{Acceleration of Computations Based on the Reuse of Search Information}
 
-As noted earlier, the MCOlex problem at each stage of the successive concessions method (\ref{eq:7}) is a global optimization problem with non-convex constraints (\ref{eq:13}).
-Generally, the solution to such problems must start at the beginning, and multiple solutions of such global optimization problems may require extensive calculations.
-However, the presence of search information $A_k$ from (\ref{eq:18}) allows us to bring the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:7}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$,  of the initial problem (\ref{eq:1}) \cite{c30,c31}.
-This re-use of search information, to find the next efficient decision, reduces the amount of computations needed to solve each next optimization problem to just a few iterations; see Section \ref{sec:4} for the results of computational experiments.
+As noted earlier, the MCOlex problem at each stage (\ref{eq:9}) of the $\varepsilon$-constraint method is a global optimization problem with nonlinear constraints (\ref{eq:16}).
+Generally, the solution to such problems should start from the very beginning; and multiple solutions to such global optimization problems may require extensive calculations.
+However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}.
+This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of computational experiments.
 
-Moreover, using search information $A_k$ allows to reduce the multistage solution of the MCO problem in accordance with the scheme (\ref{eq:7}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:7}), namely
-\begin{equation}\label{eq:26}
-P_{\delta lex} (f,\delta,D)=Arg \min_{y\in D}{f_s(y)}, f_j(y)\leq f_j^*+\delta_j, 1 \leq j < s,
+Moreover, using the search information $A_k$ allows reducing the multistage solution of the MCO problem in accordance with the scheme (\ref{eq:9}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:9}), namely
+\begin{equation}\label{eq:30}
+P_{\delta lex} (f,\delta,D)=Arg \min_{y \in D} {f_s(y)}, f_j(y)\leq f_j^* + \delta_j, 1 \leq j < s,
 \end{equation}
-where for the unknown a priori values $f_j^*$, $1 \leq j < s$, can be used the estimates $z_j^k$, $1 \leq j < s$, from (\ref{eq:20}) obtained from the search information $A_k$,  and for the constants $\delta_j$, $1 \leq j<s$, we can use the values normalized to the range of the constraints' values, that is
-\begin{equation}\label{eq:27}
+where for the unknown a priori values $f_j^*$, $1 \leq j < s$, the estimates $z_j^k$, ${1 \leq j < s}$, from (\ref{eq:23}) obtained from the search information $A_k$ can be used, and for the concessions $\delta_j$, $1 \leq j < s$, the normalized values can be applied, that is
+\begin{equation}\label{eq:31}
 	\begin{matrix}
-	\delta_j=z_j^k+\theta_j (g_j^k-z_j^k ),\; 0\leq \theta_j\leq 1, \; 1 \leq j < s,\\
-	g_j^k= \max{\{ g_j (x_i ) : \nu(x_i)=j,\; 1\leq i\leq k \}}.
+	\delta_j = z_j^k + \theta_j (g_j^k-z_j^k ), 0 \leq \theta_j \leq 1, 1 \leq j < s,\\
+	g_j^k = \max{\{ g_j (x_i ) : \nu(x_i)=j, 1 \leq i \leq k \}}.
 	\end{matrix}
@@
 
-The applicability of such a one-step scheme (\ref{eq:26})--(\ref{eq:27}) instead of the scheme (\ref{eq:7}) is computationally evaluated in Section \ref{sec:4}.
+In particular, a similar one-step scheme (\ref{eq:30})--(\ref{eq:31}) can be applied through the use of AGCS.
 
 With a complete scheme for calculating criteria values, i.e.
-when at the iteration points $x_i$, $1 \leq i \leq k$,  the values of all criteria $f_j (x_i)$, $1 \leq j \leq s$, $1 \leq i \leq k$, are calculated, the effect of using the search information $A_k$ can be even more significant.
-In this case, the search information can also be used when changing the values of the concessions $\delta_j$, $1 \leq j < s$, and thus, the solution of the next problem from the $\mathbb{P}_t$ family (\ref{eq:8}) can be carried out repeatedly using the results of all previously performed calculations.
-Results of computational experiments (see Section \ref{sec:4}), demonstrate that the volume of computations performed, based on reusing search information, is reduced by a factor of no less than 6.
+when at the iteration points $x_i$, $1 \leq i \leq k$ the values of all criteria $f_j(x_i)$, $1 \leq j \leq s$, $1 \leq i \leq k$, are calculated, the effect of using the search information $A_k$ can be even more significant.
+In this case, the search information can also be used when changing the values of the concessions $\delta_j$, $1 \leq j < s$ and thus, the solution of the next problem from the $\mathbb{P}_t$ family (\ref{eq:10}) can be carried out repeatedly using the results of all previously performed calculations.
+The results of computational experiments (see Section \ref{sec:5}), demonstrate that the volume of computations performed, based on reusing search information, can significantly decrease.
 
-The AGCS algorithm, supplemented by the ability to use search information to solve multiple MCOlex  problems, will be referred to below as the global search algorithm for multistage solutions of a set of multiextremal optimization problems with non-convex constraints (MAGCS).
+The AGCS algorithm, supplemented by the ability to use search information to solve multiple MCOlex problems, will be referred to below as the global search algorithm for multistage solutions of a set of multiextremal optimization problems with nonlinear constraints (MAGCS).
 
-\section{Results of computational experiments}
+\section{Evaluation of the Effectiveness of Search Information Reuse}
 \label{sec:4}
 
-The computational experiments were carried out on the Lobachevsky supercomputer at Nizhny Novgorod State University (the operating system CentOS 6.4, the management system SLURM).
+Evaluation of the effectiveness of multiple uses of search information is based on the key properties of the AGCS algorithm.
+When searching for the global minimum, this algorithm constructs a grid of search trial points covering the search domains.
+In addition, this algorithm is stable, i.e., in minor variations of the minimized function, the value of change in the evaluation of the global minimum is limited.
+
+To reduce the complexity of theoretical analysis, all further studies have been performed for the one-dimensional version of the AGCS algorithm; due to the use of dimensional reduction, the results obtained will also be applicable when solving multidimensional optimization problems.
+
+The initial theoretical analysis of global search properties when using search information was performed in \cite{c48}.
+The obtained theoretical statements can be reformulated as applied to lexicographic multicriteria optimization problems.
+
+\begin{theorem}
+Let the next solved optimization problem $\psi(x)$ of the computational scheme (\ref{eq:9}) differ from the previously solved problem $\phi(x)$ by some limited value $\delta$, i.e.
+\begin{equation}\label{eq:32}
+|\psi(x)-\phi(x)| \leq \delta, x \in [0,1]
+\end{equation}
+and let the stopping criterion of (\ref{eq:28}) for the specified accuracy of global search $\varepsilon>0$ be fulfilled when minimizing the function $\phi(x)$. Then 
+\begin{equation}\label{eq:33}
+|\psi(x^*)-\phi(x_k^*)| \leq \Delta = \frac{\alpha L \varepsilon}{2} + \delta,
+\end{equation}
+if starting from a certain iteration $k_0>0$ the condition is fulfilled for the estimation of the Lipschitz constant $m$ and some $\alpha > 1$
+\begin{equation}\label{eq:34}
+m \geq L (1+ \sqrt{\frac{\alpha + 4 \beta (1 + \beta)}{\alpha - 1}}),
+\end{equation}
+where
+\begin{equation}\label{eq:35}
+m=\max{r_l h_l^k}, 1 \leq l \leq M^k, k>k_0,
+\end{equation}
+\begin{equation}\label{eq:36}
+L=\max{L_i}, 1 \leq i \leq s, L_i \text{from} (\ref{eq:3}),
+\end{equation}
+\begin{equation}\label{eq:37}
+\Psi(x^*)=\min{\psi(x)},x \in [0,1],
+\end{equation}
+\begin{equation}\label{eq:38}
+x_k^* = arg \min{\phi(x_k)}, 1 \leq i \leq k, 
+\end{equation}
+(the points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $h_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively).
+\end{theorem}
+
+This statement means that if the error $\Delta$ from (\ref{eq:33}) of determination of the minimum value of $\psi(x^*)$ is acceptable, then the minimization of the function $\psi(x)$ does not require any additional iterations of the global search -- the evaluation of the minimum value of $\psi(x^*)$ can be obtained in accordance with (\ref{eq:33}), using values $x_k$, ${1 \leq i \leq k}$, located in the search information $A_k$ from (\ref{eq:21}), obtained by minimizing the function $\phi(x)$.
+
+\begin{theorem}
+Let the optimization problem $\phi(x)$ of the computational scheme (\ref{eq:9}) be solved before the stopping criterion in (\ref{eq:28}) is fulfilled for the specified accuracy of the global search $\varepsilon > 0$ and the density of trials for the interval $[a,b] \subset [0,1]$ is reached
+\begin{equation}\label{eq:39}
+p_{ab}=\frac{n_{ab}}{b-a}
+\end{equation}
+where $n_{ab}$ is the number of trial points in the interval $[a,b]$ during the minimization of the function $\phi(x)$.
+Then, when solving the next optimization problem $\psi(x)$ using the search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, the number of trials in the same interval $[a,b]$ instead of the maximum value
+\begin{equation}\label{eq:40}
+(3m/2\Delta)(b-a), m \text{from} (\ref{eq:35}),
+\end{equation}
+will be limited by the value
+\begin{equation}\label{eq:41}
+n_{ab}' \leq (3m/2\Delta - p_{ab})(b-a)
+\end{equation}
+if the condition is met
+\begin{equation}\label{eq:42}
+\psi(x) \geq \psi(x^*)+\Delta, x \in [a,b].
+\end{equation}
+\end{theorem}
+
+Considering the estimate (\ref{eq:41}) of the maximum number of trials in the search domain subintervals $[0,1]$, the following statement may be formulated.
+
+\begin{theorem}
+If the number of trials in the interval $[a,b]\subset[0,1]$ exceeds the maximum estimate (\ref{eq:41}) when solving the optimization problem $\psi(x)$ of the computational scheme (\ref{eq:9}), then when solving the next optimization problem $\psi(x)$ using the search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, additional trials in the interval $[a,b]$ will not be performed while the condition (\ref{eq:42}) is fulfilled.
+\end{theorem}
+
+The theoretical statements formulated above allow estimating the positive effect of multiple uses of search information obtained in the process of calculations.
+If the new problem is sufficiently close to the already solved problems, the evaluation of the global search can be obtained based on the available search information without performing additional iterations of the global search (Theorem 2).
+In the extreme case, when solving a sufficient number of information-connected optimization problems, each new problem can be close to one of the previously solved problems, making the already available search information sufficient for solving this new problem.
+In other cases, when a new optimization problem is quite different from the previously solved problems, the use of search information, previously obtained in the course of calculations, allows reducing the number of trials performed in the subintervals of the search domain $[0,1]$ up to the complete absence of additional trials in these subintervals (Theorems 3-4).
+
+\section{Results of Computational Experiments}
+\label{sec:5}
+
+The computational experiments were carried out on the Lobachevsky supercomputer at the Lobachevsky State University of Nizhny Novgorod (the CentOS 6.4 operating system and the SLURM workload manager).
 A supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM processors.
-Each CPU is 8-core; that is, there are 16 CPU cores available on the node.
-The Intel C++ 14.0.2 compiler was used to obtain an executable program code.
+Each CPU is 8-core, which means 16 CPU cores are available on the node.
+The Intel C ++ 14.0.2 compiler was used to obtain an executable program code.
 The numerical experiments were performed using the Globalizer system \cite{c32}.
 
-The global optimization algorithms used within the framework of the proposed approach have proven to be efficient in the process of conducting test computational experiments and have been widely used in solving practical optimization problems (see, e.g., \cite{c33,c34}).
-The results of computational experiments in solving multicriteria optimization problems are given below.
+Evaluating the efficiency of multicriteria optimization methods is not quite a trivial question.
+While the efficiency of scalar optimization methods can be evaluated by the indicators of solution accuracy and the number of optimization iterations performed (the number of the calculated function values), solving MCO problems can lead to finding several efficient decisions or even approximating the whole Pareto set.
+There are many different approaches to this issue -- for example, efficiency indicators can be determined by the mutual domination rate \cite{c49}, the degree of approximation \cite{c50}, the overall Pareto spread \cite{c51}, etc.
+Authors of \cite{c52} propose an even wider set of 57 efficiency indicators in solving MCO problems.
+Unfortunately, it should be noted that no single approach has yet been adopted to evaluate the efficiency of the MCO methods.
 
-In the first series of computational experiments, the developed MAGCS algorithm was compared with several multicriteria optimization algorithms.
-To make a comparison, we used the test two-criteria problem proposed in \cite{c35}:
-\begin{equation}\label{eq:28}
+In this article we assessed the efficiency of the MCO methods using two widely used indicators \cite{c35,c36}:
+\begin{itemize}
+	\item The hypervolume index (HV). This indicator evaluates the approximation of the Pareto set in terms of completeness; a higher value corresponds to a more complete coverage of the Pareto set.
+	\item The distribution uniformity index (DU). This indicator evaluates the uniformity of Pareto set coverage; a lower value corresponds to a more uniform coverage of the Pareto set.
+\end{itemize}
+
+In addition, since the solution of the MCO problem is reduced to the solution of multiple scalar optimization problems, the efficiency will be evaluated by ``classical'' indicators: the solution accuracy and the number of the calculated function values.
+The latter indicator is even more important because due to the initial assumption about the time-consuming calculations of efficiency criteria values of the MCO problem, this indicator evaluates the computational complexity of solving the MCO problem.
+
+Another important issue of the efficiency evaluation consists in the choice of benchmarks, which will be used to measure the efficiency of the methods applied.
+Unfortunately, there is no unified approach to this issue either.
+In many studies, the efficiency of the developed methods is demonstrated by the example of a limited set of test optimization problems.
+Researchers repeatedly attempt to resolve this issue by forming quite a wide set of test optimization problems, which can be used to evaluate the efficiency of the MCO methods.
+For instance, in \cite{c53} authors propose a flexible toolkit for constructing well-designed MCO problems.
+In \cite{c54} a set of 24 test optimization functions is given, which are widely used for evaluating the MCO methods.
+In \cite{c55} a set of 100 test problems is considered.
+In \cite{c56} an approach is proposed, in which a set of efficiency parameters of the MCO problems is formed from the well-known scalar optimization functions.
+This allows forming an almost endless set of MCO problems.
+At the same time, the set of the MCO test problems, in which the efficiency criteria would be multiextremal functions, is rather limited.
+Also, the proposed sets usually consist of single examples of the MCO problems, which makes it difficult to obtain any reliable estimates of the method efficiency.
+
+In this paper, the efficiency of the proposed methods is evaluated by executing extensive numerical experiments for solving the wide sets of the MCO problems.
+The test problems are produced by the GKLS generator \cite{c37} that can construct an almost endless number of the MCO problems using a random mechanism.
+This generator can produce multiextremal functions with predetermined characteristics (the dimensionality, the number of local minima, the size of their attraction domains, the point of global minimum, and the value of the functions, etc.).
+Fig.
+\ref{fig:3} shows an example of the two criteria generated using GKLS.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\linewidth]{fig3}
+  \caption{Contour plots of two criteria obtained by the GKLS generator}
+  \label{fig:3}
+\end{figure}
+
+As we have previously noted, the proposed global optimization algorithms were developed in the framework of the information-statistical theory of multiextremal optimization.
+These methods have proven to be efficient in numerical experiments and have been widely used for solving practical global optimization problems (see, e.g., \cite{c17,c18,c23,c26,c27,c33,c34}).
+The results of numerical experiments in solving multicriteria optimization problems are given below.
+
+In the first series of computational experiments, the MAGCS algorithm was compared with several multicriteria optimization algorithms for solving the two-criteria problem proposed in \cite{c35}:
+\begin{equation}\label{eq:43}
 f_1 (y)=(y_1-1) y_2^2+1,f_2 (y)=y_2, 0\leq y_1,y_2 \leq 1.
@@
 
-Solving the MCO problem was understood to mean constructing a numerical approximation of the Pareto domain.
-To assess the quality of the approximation, the completeness and uniformity of the Pareto domain coverage were compared, using the following two indicators \cite{c35,c36}:
-\begin{itemize}
-	\item 	The hypervolume index (HV).
-This indicator characterizes the approximation of the Pareto domain in terms of completeness; a higher value corresponds to a more complete coverage of the Pareto domain.
-	\item The distribution uniformity index (DU). This indicator characterizes the uniformity of Pareto domain coverage; a lower value corresponds to a more uniform coverage of the Pareto domain.
-\end{itemize}
+In this experiment, five multicriteria optimization algorithms were compared: the Monte-Carlo (MC) method, the genetic algorithm SEMO from the PISA library \cite{c9,c36}, the Non-uniform coverage (NUC) method \cite{c35}, the bi-objective Lipschitz optimization (BLO) method \cite{c36}, and the MAGCS algorithm.
+The results of the solution (\ref{eq:29}) for all these methods (except for MAGCS) were obtained in \cite{c36}.
 
-In the framework of this experiment, five multicriteria optimization algorithms were compared: the Monte-Carlo (MC) method, the genetic algorithm SEMO from the PISA library \cite{c9,c36}, the Non-uniform coverage (NUC) method \cite{c35}, the bi-objective Lipschitz optimization (BLO) method \cite{c36}, and the MAGCS algorithm proposed in this paper.
-The results of solving problem (\ref{eq:28}) for all these methods (except for MAGCS) were obtained in \cite{c36}.
-
-For MAGCS, 50 problems (\ref{eq:26}) were solved for different values $\theta_j$, $1 \leq j<s$, uniformly distributed in the interval [0,1].
-The accuracy from (\ref{eq:25}) $\varepsilon=0.05$ and the reliability from (\ref{eq:22}) $r=3.0$ were used.
+For MAGCS, 50 subproblems (\ref{eq:30}) were solved for different values $\theta=\theta_1$ from (\ref{eq:31}), uniformly distributed in the interval $[0,1]$.
+The accuracy $\varepsilon=0.05$ from (\ref{eq:28}) and the reliability parameters $r_1=r_2=3.0$ from (\ref{eq:29}) were used.
 The results of the executed experiments are presented in Table \ref{tab:1}.
@@
 \centering
-\caption{The numerical results of solving the multicriteria optimization problems from (\ref{eq:28})}
+\caption{The numerical results of solving the multicriteria optimization problems from (\ref{eq:43})}
 \label{tab:1}
@@
 Number of executed iterations                                                              & 500   & 500   & 515   & 498   & \textbf{273}   \\
-\begin{tabular}[c]{@{}c@{}}Number of points in \\ Pareto domain approximation\end{tabular} & 67    & 104   & 29    & 68    & \textbf{80}    \\
+\begin{tabular}[c]{@{}c@{}}Number of points in \\ the Pareto set approximation\end{tabular} & 67    & 104   & 29    & 68    & \textbf{80}    \\
 HV index                                                                                   & 0.300 & 0.312 & 0.306 & 0.308 & \textbf{0.314} \\
@@
 
-The results from the executed experiments demonstrate that the MAGCS algorithm has a significant advantage in comparison with the considered multicriteria optimization methods, even when solving relatively simple MCO problems.
+The results of the executed experiments demonstrate that the MAGCS algorithm has a significant advantage in comparison with the considered multicriteria optimization methods, even when solving relatively simple MCO problems.
 
-In the second series of computational experiments, two-criteria two-dimensional MCO problems were solved.
-Multiextremal functions obtained using the GKLS generator \cite{c37} served as the criteria for the problem.
-During the experiments, 100 multicriteria problems of this class were solved.
-In each problem, Pareto-optimal decisions were searched for 50 different values of $\theta_j$, $1\leq j < s$, uniformly distributed in the interval [0,1]; (that is, 5,000 global optimization problems were solved).
-The results obtained were averaged over the number of MCO problems solved.
+All the subsequent numerical experiments focused on the solution of 100 two-criteria MCO problems, the efficiency criteria of which were set by the GKLS generator.
+For each MCO problem from the generated set, from 10 to 100 subproblems (\ref{eq:30}) were solved for different values $\theta=\theta_1$ from (\ref{eq:31}), uniformly distributed in the interval $[0,1]$.
+The obtained results of numerical experiments were averaged by the number of solved MCO problems.
 
-In the computational experiments, calculations are terminated when the required accuracy was achieved.
-When the process was terminated, the solution was evaluated for correctness.
-As a means of control, we compared the solution points found by MAGCS and the points calculated from the Pareto boundary, taking into account the selected values of $\theta_j$, $1 \leq j < s$.
-The method parameters were set as follows: the accuracy $\varepsilon = 0.02$ and the reliability $r_1=r_2=5.6$.
-Results of the computational experiments are presented in Table 2.
+The solution of MCO problems was carried out before the stopping criterion of (\ref{eq:28}) was fulfilled, and the MCO problems were considered solved correctly if the distance of the decisons to the Pareto domain did not exceed the given accuracy $\varepsilon$ from (\ref{eq:28}).
 
-In Table 2, the first and fifth columns indicate the average number of iterations executed by the MAGCS algorithm for solving the MCOlex problems.
-The second and sixth columns contain the percentage of completed problems.
-The third, fourth, seventh and eighth columns show the values of HV and DU indicators.
-The last column gives values showing the reduction in the number of global search iterations that were executed when solving the MCOlex problems, by reusing search information.
+In the second series of experiments, the two-criteria two-dimensional MCO problems were solved.
+The solution was carried out with the method accuracy $\varepsilon = 0.025$ and the reliability parameters $r_1 = r_2 = 5.6$.
+The results of computational experiments are presented in Table \ref{tab:2}.
 
@@
 \centering
-\caption{Results of the series of experiments to solve two-dimensional two-criteria MCO problems}
+\caption{Results of the numerical experiments to solve two-criteria two-dimensional MCO problems}
 \label{tab:2}
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{ccccccccccc}
+\resizebox{\columnwidth}{!}{%
+\begin{tabular}{cccccccccccc}
 \hline
-\multicolumn{10}{c}{Search information}                                    & Reduction \\ \cline{1-9}
-\multicolumn{4}{c}{not used}         & & \multicolumn{4}{c}{used}             &  & in \\ \cline{1-4} \cline{6-9}
-Method     & Problem & Avg   & Avg   & & Method     & Problem & Avg   & Avg   &  & iteration    \\
-iter.      & solved  & HV    & DU    & & iter.      & solved  & HV    & DU    &  & number       \\ \cline{1-4} \cline{6-9} \cline{11-11}
-41 710.1   & 98.0\%  & 33.43 & 0.173 & & 2 407.41   & 99.3\%  & 33.34 & 0.224 &  & 17.3         \\ \hline
+    & \multicolumn{10}{c}{Solving   MCO problems}                                                             &     \\ \cline{2-11}
+SP  & \multicolumn{5}{c}{without   using search information} & \multicolumn{5}{c}{using   search information} & RI  \\ \cline{2-11}
+    & PI          & SPI      & SPS       & HV       & DU     & PI        & SPI    & SPS      & HV     & DU    &     \\ \hline
+10  & 8 340,1     & 834,0    & 71,3\%    & 30,34    & 0,35   & 981,7     & 98,2   & 81,9\%   & 30,10  & 0,54  & 1,0 \\
+20  & 16 756,5    & 837,8    & 78,0\%    & 30,56    & 0,31   & 1 208,0   & 60,4   & 86,8\%   & 30,25  & 0,42  & 1,6 \\
+30  & 25 209,4    & 840,3    & 80,7\%    & 30,60    & 0,32   & 1 319,2   & 44,0   & 89,7\%   & 30,33  & 0,45  & 2,2 \\
+40  & 33 625,7    & 840,6    & 81,7\%    & 30,65    & 0,31   & 1 403,0   & 35,1   & 92,4\%   & 30,40  & 0,40  & 2,8 \\
+50  & 41 994,7    & 839,9    & 82,8\%    & 30,68    & 0,31   & 1 452,8   & 29,1   & 93,5\%   & 30,46  & 0,41  & 3,4 \\
+60  & 50 428,8    & 840,5    & 83,5\%    & 30,72    & 0,29   & 1 506,6   & 25,1   & 93,8\%   & 30,46  & 0,37  & 3,9 \\
+70  & 58 923,3    & 841,8    & 84,5\%    & 30,75    & 0,29   & 1 556,9   & 22,2   & 95,1\%   & 30,52  & 0,37  & 4,4 \\
+80  & 67 318,8    & 841,5    & 84,8\%    & 30,76    & 0,30   & 1 588,5   & 19,9   & 95,3\%   & 30,51  & 0,38  & 4,9 \\
+90  & 75 684,0    & 840,9    & 85,0\%    & 30,76    & 0,25   & 1 624,7   & 18,1   & 95,5\%   & 30,57  & 0,37  & 5,4 \\
+100 & 84 164,2    & 841,6    & 84,6\%    & 30,78    & 0,26   & 1 644,3   & 16,4   & 95,9\%   & 30,58  & 0,32  & 6,0 \\ \hline
 \end{tabular}
@@
 
-Results obtained from these experiments show that the re-use of search information can reduce the amount of calculations by 17.3 times without expending additional computational resources, and according to the average HV and DU indicators, the quality of the Pareto domain remains on average at the same level.
+The first column (SP) in the Table \ref{tab:2} contains the total number of MCO subproblems to be solved.
+The PI columns show the average number of iterations for solving one MCO problem.
+The SPI columns represent the average number of iterations for solve a single MCO subproblem.
+The SPS columns contain a fraction of MCO subproblems solved with the required accuracy.
+The HV and DU columns show the hypervolume (HV) and the distribution uniformity (DU) indices of the calculated approximation of the Pareto set, respectively.
+The last column (RI) shows the degree of the reduction in the number of global search iterations performed when solving subproblems (\ref{eq:30}) by reusing the search information.
 
-In the third series of computational experiments, the two-criteria four-dimensional MCO problems were solved.
-As in the second series of computational experiments, 100 multicriteria problems were solved.
-In each problem, the Pareto-optimal decisions were calculated for 50 different values of $\theta_j$, $1 \leq j <s$, uniformly distributed in the interval $[0,1]$.
-The criteria of the MCO problems being solved were determined using the GKLS generator \cite{c37}.
-The method parameters were set as follows: the accuracy $\varepsilon = 0.025$ and the reliability $r_1=r_2=5.6$.
-The results of computational experiments are presented in Table 3.
+The results presented in Table \ref{tab:2} demonstrate that the use of search information allows to obtain an estimate of the Pareto domain after a substantially smaller total number of iterations of the global search (the PI columns).
+It should also be noted that as the number of solved subproblems increases, the MCO allows to build estimates of the Pareto region with better HV and DU indices.
+
+For a better visual representation, Figure \ref{fig:4} shows the average number of iterations performed to solve a single MCO subproblem with and without search information.
+The presented graphs convincingly demonstrate a significant reduction in the number of global search iterations when solving the MCO subproblems using search information.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=1\linewidth]{fig4}
+  \caption{Average number of iterations required to solve one MCO subproblem depending on the total number of subproblems solved}
+  \label{fig:4}
+\end{figure}
+
+Fig.
+\ref{fig:5} shows the reduction in the number of global search iterations when solving MCO subproblems using search information.
+The graphs presented show that when the number of solved MCO subproblems grows linearly, the number of iterations required to solve one subproblem is reduced almost linearly as well.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=1\linewidth]{fig5}
+  \caption{Reduction of the number of iterations performed to solve one MCO subproblem depending on the total number of subproblems solved}
+  \label{fig:5}
+\end{figure}
+
+It should be noted that the number of iterations of the global search is also reduced in the course of sequential solution of subtasks of the same MCO task.
+
+It should be noted that the number of global search iterations is also reduced in the course of sequential solution of the subproblems of the same MCO problem.
+Table \ref{tab:3} illustrates this trend showing the average number of iterations required to find a solution to one MCO subproblem, with subdividing the sequence of solved subproblems in groups of 10 MCO subproblems.
+The first row (SP) contains the MCO subproblem numbers belonging to the individual subproblem groups.
+The PI rows show the total number of iterations of the global search performed for all previous problems.
+The rows labeled as ``PI new'' show the number of iterations performed to solve the next group of MCO subproblems with the numbers of the problems given in each column heading.
+The last row labeled as ``FPI new'' shows the proportion of new iterations of global search performed to solve the next group of MCO subproblems.
 
@@
 \centering
-\caption{Results of the series of experiments to solve two-criteria four-dimensional MCO}
+\caption{Average number of iterations to solve two-criteria two-dimensional MCO problems}
 \label{tab:3}
-\resizebox{\columnwidth}{!}{
+\resizebox{\columnwidth}{!}{%
 \begin{tabular}{ccccccccccc}
 \hline
-\multicolumn{10}{c}{Search information}                                        & Reduction   \\ \cline{1-9}
-\multicolumn{4}{c}{not used}       & & \multicolumn{4}{c}{used}                & &  in       \\ \cline{1-4} \cline{6-9}
-Method      & Problem & Avg   & Avg   & & Method     & Problem & Avg   & Avg   & & iteration \\
-iter.       & solved  & HV    & DU    & & iter.      & solved  & HV    & DU    & & number    \\ \cline{1-4} \cline{6-9} \cline{11-11}
-4 536 377.9 & 83.0\%  & 30.67 & 0.335 & & 709 014.9  & 94.5\%  & 30.46 & 0.405 & & 6.4       \\ \hline
+SP       & 1-10   & 11-20  & 21-30  & 31-40  & 41-50  & 51-60  & 61-70  & 71-80  & 81-90  & 91-100 \\ \hline
+\multicolumn{11}{c}{Solving MCO problems without using search information}                         \\ \hline
+PI       & 8868   & 17007  & 26228  & 34320  & 44117  & 53776  & 63292  & 71816  & 78824  & 84164  \\
+PI new   & 8868   & 8139   & 9221   & 8092   & 9796   & 9660   & 9515   & 8525   & 7007   & 5341   \\ \hline
+\multicolumn{11}{c}{Solving MCO problems with using search information}                            \\ \hline
+PI       & 993,1  & 1230,3 & 1309,6 & 1393,2 & 1409,5 & 1430,9 & 1466,1 & 1538,2 & 1624,9 & 1644,3 \\
+PI new   & 993,1  & 237,1  & 79,3   & 83,6   & 16,3   & 21,4   & 35,2   & 72,1   & 86,7   & 19,4   \\ \hline
+\multicolumn{11}{c}{Proportion of new iterations   for the solution of a new group of subproblems} \\ \hline
+FPI new  & -      & 23,9\% & 6,4\%  & 6,4\%  & 1,2\%  & 1,5\%  & 2,5\%  & 4,9\%  & 5,6\%  & 1,2\%  \\ \hline
 \end{tabular}
@@
 
-Results from the executed experiments show that with increased dimensions of MCO problems to be solved and a corresponding increase in the volume of calculations, the efficiency of the MAGCS algorithm remains at a high level;  its use has efficiently reduced the number of the executed iterations by 6.4 times.
+Fig.
+\ref{fig:6} shows the values of the achieved reduction in the number of additional iterations required to solve the next set of subproblems (the last row of Table \ref{tab:3}).
+The above data demonstrates that solving the next group of MCO subproblems requires about $6\%$ of new global search iterations in addition to the available search information.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=1\linewidth]{fig6}
+  \caption{Proportion of new global search iterations for solving the next group of MCO subproblems}
+  \label{fig:6}
+\end{figure}
+
+The third set of computational experiments was designed to demonstrate the scalability of the proposed approach by solving more complex two-criteria four-dimensional MCO problems under the same conditions as in the previous series of calculations.
+The results of the experiments are presented in Tables \ref{tab:4}--\ref{tab:5} and in Fig.
+\ref{fig:4}--\ref{fig:6}.
+These data demonstrate that the experiments carried out to solve more complex MCO problems confirm the conclusions made earlier -- for example, solving the next group of MCO subproblems requires, in addition to the available search information, from $0.5\%$ to $14.2\%$ more global search iterations (Table \ref{tab:5}, last row).
+
+\begin{table}[ht]
+\centering
+\caption{Results of the numerical experiments to solve two-criteria four-dimensional MCO problems}
+\label{tab:4}
+\resizebox{\columnwidth}{!}{%
+\begin{tabular}{cccccccccccc}
+\hline
+    & \multicolumn{10}{c}{Solving MCO problems}                                                                &     \\ \cline{2-11}
+SP  & \multicolumn{5}{c}{without using search information} & \multicolumn{5}{c}{With using search information} & RI  \\ \cline{2-11}
+    & PI$\times 10^3$& SPI$\times 10^3$& SP       & HV      & DU     & PI$\times 10^3$& SPI$\times 10^3$& SP       & HV      & DU    &     \\ \hline
+10  & 1 088,7    & 108,8     & 71,3\%   & 30,34   & 0,35   & 340,4    & 34,0      & 81,9\%   & 30,10   & 0,54  & 1,0 \\
+20  & 2 245,6    & 112,2     & 78,0\%   & 30,56   & 0,31   & 518,3    & 25,9      & 86,8\%   & 30,25   & 0,42  & 1,3 \\
+30  & 3 407,8    & 113,5     & 80,7\%   & 30,60   & 0,32   & 607,9    & 20,2      & 89,7\%   & 30,33   & 0,45  & 1,7 \\
+40  & 4 525,3    & 113,1     & 81,7\%   & 30,65   & 0,31   & 669,6    & 16,7      & 92,4\%   & 30,40   & 0,40  & 2,0 \\
+50  & 5 677,0    & 113,5     & 82,8\%   & 30,68   & 0,31   & 717,2    & 14,3      & 93,5\%   & 30,46   & 0,41  & 2,4 \\
+60  & 6 854,4    & 114,2     & 83,5\%   & 30,72   & 0,29   & 750,3    & 12,5      & 93,8\%   & 30,46   & 0,37  & 2,7 \\
+70  & 7 965,3    & 113,7     & 84,5\%   & 30,75   & 0,29   & 782,3    & 11,1      & 95,1\%   & 30,52   & 0,37  & 3,0 \\
+80  & 9 134,8    & 114,1     & 84,8\%   & 30,76   & 0,30   & 807,4    & 10,0      & 95,3\%   & 30,51   & 0,38  & 3,4 \\
+90  & 10 280,2   & 114,2     & 85,0\%   & 30,76   & 0,25   & 822,7    & 9,1       & 95,5\%   & 30,57   & 0,37  & 3,7 \\
+100 & 11 401,9   & 114,0     & 84,6\%   & 30,78   & 0,26   & 836,3    & 8,3       & 95,9\%   & 30,58   & 0,32  & 4,1 \\ \hline
+\end{tabular}
+}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Results of the numerical experiments to solve two-criteria four-dimensional MCO problems}
+\label{tab:5}
+\resizebox{\columnwidth}{!}{%
+\begin{tabular}{ccccccccccc}
+\hline
+SP          & 1-10   & 11-20  & 21-30  & 31-40  & 41-50  & 51-60  & 61-70  & 71-80  & 81-90   & 91-100  \\ \hline
+\multicolumn{11}{c}{Solving MCO problems without using   search information}                            \\ \hline
+PI$\times 10^3$& 1167,1 & 2320,4 & 3403,4 & 4530,1 & 5543,4 & 6603,5 & 7968,0 & 9614,4 & 10832,6 & 11402,0 \\
+PI new$\times 10^3$& 1167,1 & 1153,3 & 1083,0 & 1126,7 & 1013,3 & 1060,1 & 1364,5 & 1646,4 & 1218,2  & 569,4   \\ \hline
+\multicolumn{11}{c}{Solving MCO problems with using   search information}                               \\ \hline
+PI$\times 10^3$& 399,5  & 556,1  & 635,1  & 686,1  & 690,3  & 714,4  & 769,4  & 837,3  & 886,7   & 890,7   \\
+PI new$\times 10^3$& 399,5  & 156,6  & 79,0   & 51,0   & 4,2    & 24,2   & 54,9   & 67,9   & 49,4    & 4,0     \\ \hline
+\multicolumn{11}{c}{Proportion   of new iterations required to solve the next group of subproblems}     \\ \hline
+FPI new  & -      & 39,2\% & 14,2\% & 8,0\%  & 0,6\%  & 3,5\%  & 7,7\%  & 8,8\%  & 5,9\%   & 0,5\%   \\ \hline
+\end{tabular}
+}
+\end{table}
 
 \section{Conclusion}
-\label{sec:5}
+\label{sec:6}
 
-This paper proposes a new approach for solving computationally expensive lexicographic multicriteria optimization problems (MCOlex), in which the efficiency criteria can be multiextremal, and calculating the criteria values may require a large amount of computation.
-A key feature of this class of problems is the ability, during the computation process, to alter the ordering of efficiency criteria in terms of importance, which in turn necessitates a multistage solution for MCOlex problems.
+In this paper a new approach is proposed for solving computationally expensive lexicographic multicriteria optimization problems (MCOlex), which can be complicated by the multiextremal nature of the efficiency criteria and extensive volume of computations required to calculate the criteria values.
+A key feature of this class of problems is the ability to alter the ranking of efficiency criteria in terms of importance during calculations, which in turn necessitates a multistage solution for MCOlex problems.
 
-Overcoming the enormous computational complexity of addressing the formulated new class of MCOlex problems is ensured by solving a sequence of global optimization problems with non-convex constraints using efficient information-statistical methods of global optimization by using an original index constraint accounting scheme instead of the commonly used penalty functions.
-A core element in this approach is the ability to use all the search information obtained in the computation process, utilizing the multistage solution of MCOlex problems.
-Starting to solve each new stage of the solution, this search information allows us to incorporate the previously calculated values of efficiency criteria into the values of the next scalar problem of multiextremal optimization.
-The search information retrieved becomes part of the current optimization state and through optimization methods, results in adaptive planning of the iterations to run the global search.
+The computational complexity of addressing the formulated new class of MCOlex problems can be overcome by solving a sequence of global optimization problems with nonlinear constraints using efficient information-statistical methods of global optimization.
+A core element in this approach is the ability to use all the search information obtained in the course of solving the multistage MCOlex problems.
+With each new stage of the solution, this search information allows incorporating the previously calculated values of efficiency criteria into the values of the next scalar problem of multiextremal optimization.
+The search information retrieved becomes part of the current optimization state and, through optimization methods, results in adaptive planning of the iterations to run the global search.
 
-Results of computational experiments demonstrate that this approach can significantly reduce the computational complexity of multistage MCOlex problem solving.
+The results of numerical experiments demonstrate that this approach can sufficiently reduce the computational complexity of multistage MCOlex problem-solving.
 
-In conclusion, it must be noted that the approach described here is promising, and requires further research.
-Importantly, it is necessary to continue carrying out computational experiments to solve multicriteria optimization problems with a larger number of efficiency criteria and for greater dimensions in the optimization problems to be solved.
+In conclusion, it can be noted that the approach described here is promising, and requires further study.
+It is necessary to continue numerical experiments to solve multicriteria optimization problems with a larger number of efficiency criteria and for greater dimensions in the optimization problems.
 It is also necessary to evaluate the possibility of parallel computing using high-performance supercomputer systems.

### 2020-06-22 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
fix institute name

@@
 \usepackage{amsmath}
+
 %\usepackage{hyperref}
@@
 \institute{               
-V. Gergel \at
-Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
+Victor Gergel \at
+%\at Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
 \email{gergel@unn.ru} 
 \and
-E. Kozinov \at
-Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
+Evgeniy Kozinov \at
+%\at Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
 \email{evgeny.kozinov@itmm.unn.ru} 
 \and
-K. Barkalov \at
-Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
-\email{konstantin.barkalov@itmm.unn.ru}
+Konstantin Barkalov \at
+%\at Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
+\email{konstantin.barkalov@itmm.unn.ru} \\
+\at Institute of Informational Technologies, Mathematics and Mechanics, Mathematical Center, Lobachevsky State University of Nizhni Novgorod, Gagarin ave., 23, Nizhni Novgorod, Russia 
 }

### 2020-06-26 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
OptLet fix

@@
 Finding the whole set of efficient decisions may require a lot of computations and is often deemed redundant for the decision-maker.
-As a result, approaches that allow obtaining only particular limited sets of effective decisions are used in practice more often.
+As a result, approaches that allow obtaining only particular limited sets of efficient decisions are used in practice more often.
 
@@
 When criteria are ranked by importance, an MCO problem can be solved via consecutive optimization of criteria in the order of their importance: the most important criterion is optimized first, followed by the optimization of the second criterion based on the analysis of the set of decisions, in which the first criterion reaches the best value.
-This sequence is simple and clear and allows using many effective scalar optimization algorithms.
+This sequence is simple and clear and allows using many efficient scalar optimization algorithms.
 Such a multistage sequence can be reduced to solving one scalar optimization problem by linear convolution of criteria \cite{c38}.
 At the same time, this approach excessively reduces the number of calculated efficient decisions, which limits the number of options for the decision-maker.
-The $\varepsilon$-constraint scalarization approach can be used to expand the set of effective decisions that can be considered by the decision-maker.
+The $\varepsilon$-constraint scalarization approach can be used to expand the set of efficient decisions that can be considered by the decision-maker.
 In this case, only the most important criterion is used for optimization, while the maximum permissible value is set for other criteria (i.e., all other criteria are transformed into MCO problem constraints) \cite{c39,c40}.
-Some researchers \cite{c41} describe the extended version of the $\varepsilon$-constraint method, in which violation of restrictions is allowed if it leads to a noticeable improvement of values by certain criteria of an MCO problem.
+Some researchers \cite{c41} describe the extended version of the $\varepsilon$-constraint method, in which violation of constraints is allowed if it leads to a noticeable improvement of values by certain criteria of an MCO problem.
 Application of this approach is rather complex because the maximum allowable values (the concessions) of criteria must be set initially before calculations begin, which often causes substantial difficulties.
 A proposed solution \cite{c42} of this issue is to combine the initial solution sequence for lexicographical optimization problems, when concessions can be set in the process of calculations after each subsequent optimization stage (although such a scheme again requires the solution of the series of scalar optimization problems).
-The approach proposed in \cite{c43} allows even greater freedom in formulating problems of searching for effective decisions, when the order of criteria importance can also change dynamically.
-It is necessary to note that the presence of variability in problem statement for the search of effective decisions is very important as requirements to optimality of selected decisions can be varied by the decision-maker in the course of calculations.
+The approach proposed in \cite{c43} allows even greater freedom in formulating problems of searching for efficient decisions, when the order of criteria importance can also change dynamically.
+It is necessary to note that the presence of variability in problem statement for the search of efficient decisions is very important as requirements to optimality of selected decisions can be varied by the decision-maker in the course of calculations.
 Such variability is also provided in other approaches to solving MCO problems by various methods of criteria scalarization -- see, for example, \cite{c31,c44}.
@@
 
-Such unimprovable decisions are called Pareto efficient or optimal.
+Such unimprovable decisions are called Pareto efficient or Pareto optimal.
 
@@
 f^*_i =\min_{y \in D}{f_i(y)},  f_j(y) \leq f^*_j + \delta_j, 1 \leq j < i, 1 < i \leq s, \\
-P_{lex}^\delta(f,\delta,D) = Arg \min_{y \in D}{f_s(y)},f_j(y) \leq f^*_j + \delta_j, 1 \leq j < i, 1 < i \leq s.
+P_{lex}^\delta(f,\delta,D) = Arg \min_{y \in D}{f_s(y)},f_j(y) \leq f^*_j + \delta_j, 1 \leq j < s.
 \end{matrix}
@@
 \begin{itemize}
-	\item Since the estimate of the global minimum is an integral characteristic of the optimization problem (the value of the minimized function in the point of the global minimum should be compared with the values of the function in all other points of the search domain), the global search methods should provide at least some degree of coverage of the search domain,
+	\item Since the estimate of the global minimum is an integral characteristic of the optimization problem (the value of the minimized function at the point of the global minimum should be compared with the values of the function at all other points of the search domain), the global search methods should provide at least some degree of coverage of the search domain,
 	\item Due to the ``curse of dimensionality'' reducing computational complexity requires the resulting coverage of the search domain to be heterogeneous (nonuniform), i.e.
 it must be dense only in the vicinity of globally optimal points,
-	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- on eof the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}),
-	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function in these points), which significantly increases the complexity and computational cost of the global optimization methods.
+	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- on one of the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}),
+	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function at these points), which significantly increases the complexity and computational cost of the global optimization methods.
 Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47}, the Peano space-filling curves \cite{c17,c23},
@@
 	\item Using the global optimization methods that generate nonuniform coverages of the search domain;
-	\item Intensive search information obtained in the process of computations for solving multiple global optimization problems.
+	\item Intensive use of the search information obtained in the process of computations for solving multiple global optimization problems.
 \end{itemize}
@@
 
-The nonuniform coverage of the search domain $D$ can be constructed adaptively in the following manner: when choosing points for the next optimization iterations, the global search is performed taking into account all the available search information (the points of previous search iterations and the values of the optimized function in these points), that is
+The nonuniform coverage of the search domain $D$ can be constructed adaptively in the following manner: when choosing points for the next optimization iterations, the global search is performed taking into account all the available search information (the points of previous search iterations and the values of the optimized function at these points), that is
 \begin{equation}\label{eq:11}
@@
 \begin{equation}\label{eq:24}
-h^k_l=\max{\{|z_i-z_j|/\sqrt[N]{x_i-x_j}, \nu_i=\nu_j=l, i>j, 1\leq l\leq M^k\}}.
+H^k_l=\max{\{|z_i-z_j|/\sqrt[N]{x_i-x_j}, \nu_i=\nu_j=l, i>j, 1\leq l\leq M^k\}}.
 \end{equation}
 
-If $h_l^k$ in (\ref{eq:24}) turns out to be indefinite or equal to zero, then $h_l^k=1$ is taken.
+If $H_l^k$ in (\ref{eq:24}) turns out to be indefinite or equal to zero, then $H_l^k=1$ is taken.
 For the initial values of these parameters, $M^0 = H_l^0 = 1$, $z_l^0 = 0$, $1 \leq i \leq m+1$, can be taken.
@@
 	\item the functions $g_j$, $1 \leq j \leq m+1$, $[0,1]$ satisfy the H\"older condition of (\ref{eq:13}) with the constants $H_j$, $1 \leq j \leq m+1$,
-	\item for the values $h_\nu$ from (\ref{eq:24}), starting from a certain step, the inequalities are valid\footnote{The values $r_{\nu} > 1$, $1 \leq \nu \leq m+1$,are the AGCS reliability parameters used for computing the interval characteristics in (\ref{eq:25}).}
+	\item for the values $H^k_\nu$ from (\ref{eq:24}), starting from a certain step, the inequalities are valid\footnote{The values $r_{\nu} > 1$, $1 \leq \nu \leq m+1$,are the AGCS reliability parameters used for computing the interval characteristics in (\ref{eq:25}).}
 \begin{equation} \label{eq:29}
-r_\nu h_\nu > 2H_\nu, r_\nu > 1, 1 \leq \nu \leq m+1,
+r_\nu H^k_\nu > 2H_\nu, r_\nu > 1, 1 \leq \nu \leq m+1,
 \end{equation}
@@
 However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}.
-This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of computational experiments.
+This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of numerical experiments.
 
@@
 In this case, the search information can also be used when changing the values of the concessions $\delta_j$, $1 \leq j < s$ and thus, the solution of the next problem from the $\mathbb{P}_t$ family (\ref{eq:10}) can be carried out repeatedly using the results of all previously performed calculations.
-The results of computational experiments (see Section \ref{sec:5}), demonstrate that the volume of computations performed, based on reusing search information, can significantly decrease.
+The results of numerical experiments (see Section \ref{sec:5}), demonstrate that the volume of computations performed, based on reusing search information, can significantly decrease.
 
@@
 \begin{equation}\label{eq:35}
-m=\max{r_l h_l^k}, 1 \leq l \leq M^k, k>k_0,
+m=\max{r_l H_l^k}, 1 \leq l \leq M^k, k>k_0,
 \end{equation}
 \begin{equation}\label{eq:36}
-L=\max{L_i}, 1 \leq i \leq s, L_i \text{from} (\ref{eq:3}),
+L=\max{L_i}, 1 \leq i \leq s, L_i\; \text{from}\; (\ref{eq:3}),
 \end{equation}
 \begin{equation}\label{eq:37}
-\Psi(x^*)=\min{\psi(x)},x \in [0,1],
+\psi(x^*)=\min{\psi(x)},x \in [0,1],
 \end{equation}
@@
 \end{equation}
-(the points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $h_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively).
-\end{theorem}
+\end{theorem}(the points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $H_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively).
 
@@
 \begin{equation}\label{eq:40}
-(3m/2\Delta)(b-a), m \text{from} (\ref{eq:35}),
+(3m/2\Delta)(b-a), m \; \text{from} \;  (\ref{eq:35}),
 \end{equation}
@@
 
-\section{Results of Computational Experiments}
+\section{Results of Numerical Experiments}
 \label{sec:5}
 
-The computational experiments were carried out on the Lobachevsky supercomputer at the Lobachevsky State University of Nizhny Novgorod (the CentOS 6.4 operating system and the SLURM workload manager).
+The numerical experiments were carried out on the Lobachevsky supercomputer at the Lobachevsky State University of Nizhny Novgorod (the CentOS 6.4 operating system and the SLURM workload manager).
 A supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM processors.
@@
 
-In this article we assessed the efficiency of the MCO methods using two widely used indicators \cite{c35,c36}:
+In this paper we assessed the efficiency of the MCO methods using two widely used indicators \cite{c35,c36}:
 \begin{itemize}
@@
 In \cite{c55} a set of 100 test problems is considered.
-In \cite{c56} an approach is proposed, in which a set of efficiency parameters of the MCO problems is formed from the well-known scalar optimization functions.
+In \cite{c56} an approach is proposed, in which a set of efficiency criteria of the MCO problems is formed from the well-known scalar optimization functions.
 This allows forming an almost endless set of MCO problems.
@@
 
-In the first series of computational experiments, the MAGCS algorithm was compared with several multicriteria optimization algorithms for solving the two-criteria problem proposed in \cite{c35}:
+In the first series of numerical experiments, the MAGCS algorithm was compared with several multicriteria optimization algorithms for solving the two-criteria problem proposed in \cite{c35}:
 \begin{equation}\label{eq:43}
@@
 
-The solution of MCO problems was carried out before the stopping criterion of (\ref{eq:28}) was fulfilled, and the MCO problems were considered solved correctly if the distance of the decisons to the Pareto domain did not exceed the given accuracy $\varepsilon$ from (\ref{eq:28}).
+The solution of MCO problems was carried out before the stopping criterion of (\ref{eq:28}) was fulfilled, and the MCO problems were considered solved correctly if the distance of the decisons to the Pareto set did not exceed the given accuracy $\varepsilon$ from (\ref{eq:28}).
 
@@
 The solution was carried out with the method accuracy $\varepsilon = 0.025$ and the reliability parameters $r_1 = r_2 = 5.6$.
-The results of computational experiments are presented in Table \ref{tab:2}.
+The results of numerical experiments are presented in Table \ref{tab:2}.
 
@@
 
-The results presented in Table \ref{tab:2} demonstrate that the use of search information allows to obtain an estimate of the Pareto domain after a substantially smaller total number of iterations of the global search (the PI columns).
-It should also be noted that as the number of solved subproblems increases, the MCO allows to build estimates of the Pareto region with better HV and DU indices.
+The results presented in Table \ref{tab:2} demonstrate that the use of search information allows to obtain an estimate of the Pareto set after a substantially smaller total number of iterations of the global search (the PI columns).
+It should also be noted that as the number of solved subproblems increases, the MCO allows to build estimates of the Pareto set with better HV and DU indices.
 
-For a better visual representation, Figure \ref{fig:4} shows the average number of iterations performed to solve a single MCO subproblem with and without search information.
+For a better visual representation, Fig.
+\ref{fig:4} shows the average number of iterations performed to solve a single MCO subproblem with and without search information.
 The presented graphs convincingly demonstrate a significant reduction in the number of global search iterations when solving the MCO subproblems using search information.
@@
 
-It should be noted that the number of iterations of the global search is also reduced in the course of sequential solution of subtasks of the same MCO task.
+It should be noted that the number of iterations of the global search is also reduced in the course of sequential solution of subproblems of the same MCO problems.
 
@@
 
-The third set of computational experiments was designed to demonstrate the scalability of the proposed approach by solving more complex two-criteria four-dimensional MCO problems under the same conditions as in the previous series of calculations.
+The third set of numerical experiments was designed to demonstrate the scalability of the proposed approach by solving more complex two-criteria four-dimensional MCO problems under the same conditions as in the previous series of calculations.
 The results of the experiments are presented in Tables \ref{tab:4}--\ref{tab:5} and in Fig.
@@
 
-In conclusion, it can be noted that the approach described here is promising, and requires further study.
+In conclusion, it can be noted that the approach described here is promising, and requires further studies.
 It is necessary to continue numerical experiments to solve multicriteria optimization problems with a larger number of efficiency criteria and for greater dimensions in the optimization problems.
@@
 pp 31-40 (2010)
-\bibitem{c44} Charles Audet, Gilles Savard, and Walid Zghal: Multiobjective Optimization Through a Series of Single-Objective Formulations SIAM J. Optim.,19(1). pp. 188--210 (2008)
+\bibitem{c44} Audet, C., Savard, G., Zghal, W.: Multiobjective Optimization Through a Series of Single-Objective Formulations SIAM J. Optim.,19(1). pp. 188--210 (2008)
 \bibitem{c45} Sergeyev, Y.D., Kvasov, D.E.: Deterministic Global Optimization: An Introduction to the Diagonal Approach (SpringerBriefs in Optimization). Springer. (2017)
@@
 \bibitem{c55} Cust\'odio, A.L., Aguilar, J.F. Madeira, Ismael Vaz, Lu\'ls N. Vicente: Direct Multisearch for Multiobjective Optimization. IAM Journal on Optimization 21(3). pp. 1109-1140 (2011)
-\bibitem{c56} Dimo Brockhoff, Tea Tusar, Anne Auger, Nikolaus Hansen: Using Well-Understood Single-Objective Functions in Multiobjective Black-Box Optimization Test Suites.
+\bibitem{c56} Brockhoff, D., Tusar, T., Auger, A., Hansen, N.: Using Well-Understood Single-Objective Functions in Multiobjective Black-Box Optimization Test Suites.
 (2019) https://arxiv.org/abs/1604.00359

### 2020-06-26 2020/Saransk_Usova/theses.tex
Thesis No 1 to Saransk - first version

% !!!!!!!!!!!!!!!!!!   ВНИМАНИЕ !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
% Заголовки разделов формируются при помощи команд \section{}, \subsection{}, \subsubsection{}
% Не используйте уровень вложенности заголовков больше трех!
% -----------------------------
% Для оформления теорем, лемм, следствий используйте окружения 
% Def     - Определение
% Teor    - Теорема
% Lem     - Лемма
% Predl   - Предложение
% Ass     - Утверждение
% Cor     - Следствие
% Example - Пример
% -----------------------------
% Доказательство теоремы начинается командой \proof и завершается командой \endproof
% -----------------------------
% Литература помещается в окружение biblio.

\documentclass[11pt, oneside, a4paper]{article}
%\usepackage[cp1251]{inputenc} % кодировка
\usepackage[utf8]{inputenc} % кодировка
\usepackage[english, russian]{babel} % Русские и английские переносы
\usepackage{graphicx}          % для включения графических изображений
\usepackage{cite}              % для корректного оформления литературы
\usepackage{enumitem}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{mathtext}

%стилевой пакет
\usepackage{schoolseminar2020}

%\renewcommand{\thefootnote}{*}\footnote{Работа выполнена при частичной финансовой поддержке .....}

\begin{document}
% \udk     - универсальный десятичный классификатор
% \msc     - Индекс предметной классификации (Mathematics Subject Classification)
% \title   - название статьи
% \authors - список авторов
\setcounter{page}{1}

\udk{519.853.4}

\title{
%Способ учета наличия разрывов в задачах многоэкстремальной оптимизации 
Решение задач глобальной оптимизации с учетом возможных разрывов целевой функции
\footnote{Исследование выполнено при финансовой поддержке РФФИ в рамках научного проекта № 19-07-00242.}
}

\authors{К.А. Баркалов, М.А. Усова}
\organizations{Нижегородский государственный университет им. Н.И. Лобачевского}

% \section{название} - заголовок раздела первого уровня
% \subsection{название} - заголовок раздела второго уровня
% \subsubsection{название} - заголовок раздела третьего уровня

\bigskip

В работе рассматривается задача поиска глобального минимума $x^*$ одномерной функции $\varphi(x)$ вида
\begin{equation}\label{problem}
\varphi(x^*)=\min\left\{\varphi(x):x\in\left[a,b\right]\right\}.
\end{equation}
Предполагается, что целевая функция является многоэкстремальной и задана в виде ``черного ящика'', т.е. некоторого алгоритма вычисления значения функции в зависимости от параметра. При этом каждое \textit{испытание} (т.е. вычисление значения функции в области поиска) является трудоемкой операцией.

Задачи одномерной оптимизации играют важную роль, т.к. многие подходы к решению многомерных задач оптимизации так или иначе основаны на сведении решения исходной многомерной задачи к решению серии связанных задач одномерной оптимизации (см., например, \cite{Grishagin2007,Sergeyev2008}).
В литературе описаны различные методы решения задачи (\ref{problem}) в зависимости от предположений о свойствах целевой функции. Одним из подобных предположений, которое часто выполняется в прикладных оптимизационных задачах, является предположение о том, что целевая функция $\varphi(x)$ удовлетворяет условию Липшица с априори неизвестной константой $L$
\[
\left|\varphi(x')-\varphi(x'')\right|\leq L\left|x'-x''\right|,\; x',x'' \in [a,b],\; 0<L<\infty.
\]
Использование данного свойства целевой функции является типичным для многих подходов к разработке алгоритмов многоэкстремальной оптимизации \cite{Evtushenko2009,Elsakov}.

Однако в некоторых прикладных задачах свойство липшицевости может не выполняться в силу наличия скачкообразных изменений значений функции в ряде точек области поиска. Мы будем интерпретировать подобные резкие изменения как разрывы первого рода. Иногда множество точек, в которых характеристики объекта претерпевают скачок, известно заранее. Вместе с тем существуют задачи, в которых нет априорных оценок точек разрыва, но есть информация, что такие точки возможны.

Известные методы решения подобных задач, как правило, либо обобщают понятие градиента для разрывных функций \cite{Batuhtin1997}, либо принадлежат классу биоинспирированных алгоритмов \cite{Jihui}. Указанные методы обеспечивают, вообще говоря, поиск только \textit{локального решения} задачи.

В докладе на конференции будет представлен разработанный в ННГУ им. Н.И. Лобачевского эффективный алгоритм для поиска \textit{глобального решения} многоэкстремальных задач с разрывными функциями. 
Данный алгоритм является дальнейшим развитием \textit{алгоритма глобального поиска}, описание, теория сходимости и различные модификации которого представлены в \cite{Strongin2013}.

В новом алгоритме глобального поиска для разрывных функций (АГП-Р) учет разрывов целевой функции происходит явно, т.е. без операции сглаживания разрыва. 
При этом охватываются случаи как заданных, так и не заданных точек разрыва: при отсутствии информации о координатах точек разрыва их оценка проводится адаптивно, на основе вычисленных значений функции в точках проведенных испытаний.  
Ошибочное определение некоторых точек разрыва не приводит к потере сходимости метода. Решающие правила алгоритма интерпретируют такие точки как случай устранимого разрыва, поскольку левый и правый пределы оптимизируемой функции в таких точках совпадают.

%Более подробный комментарий?

В качестве примера рассмотрим задачу минимизации функции
\begin{equation}\label{test}
\varphi(x) = 0.1 \sum_{i=1}^5{i\sin(10(i+1)x+i)}+
\begin{cases}
	-4,  &0   \leq x < 3.1\\
	10,  &3.1 \leq x < 4.6\\
	 0,  &4.6 \leq x < 7.0\\
	30,  &7.0 \leq x < 9.0\\
	 0,  &9.0 \leq x \leq 10.0
\end{cases}
\end{equation}
при $x\in[0,10]$.

Для решения задачи использовался алгоритм глобального поиска для разрывных функций (АГП-Р), точность поиска решения была выбрана $\epsilon = 0.001$. На рис. \ref{ris1} изображен график функции, на котором наглядно видны точки разрывов. 
Штрихами под графиком обозначены точки поисковых испытаний, потребовавшихся алгоритму для решения задачи с указанной точностью. 
В случае, когда точки разрывов были известны, метод провел XXX испытаний (штрихи верхнего ряда). В случае, когда точки разрывов не были заданы заранее и идентифицировались в процессе поиска, методу потребовалось YYY испытаний (штрихи нижнего ряда). В обоих случаях распределение точек испытаний показывает наличие сходимости лишь к глобальному экстремуму, накопления точек около локальных экстремумов --- отсутствуют.

\begin{figure}%[!ht]
	\begin{center}
			\includegraphics[width=0.7\linewidth]{ris1.png}
			\caption{График разрывной целевой функции и точки испытаний} %% подпись к рисунку
      \label{ris1}
	\end{center}
\end{figure}

Для демонстрации эффективности алгоритма глобального поиска для разрывных функций (АГП-Р), проведем его сравнение с алгоритмом имитации отжига (ИО), реализованным в MATLAB Global Optimization Toolbox. В качестве основного критерия сравнения будем использовать число поисковых испытаний (т.е. вычислений значений целевой функции), выполненное методом до достижения сходимости.

Сравнение алгоритмов проводилось при решении 100 задач с разрывной целевой функцией, которые были построены аналогично задаче (\ref{test}). В таблице 1 представлено число решенных задач и среднее число испытаний, выполненных методами при решении серии. Задача считалась решенной, если для точки некоторого испытания $x^k$ выполнялось условие $\left|x^k-x^*\right|\leq \delta$, где $x^*$ --- известное решение задачи, а $\delta = 10^{-2}$. При этом точность, используемая в условии остановки методов, была выбрана на порядок меньшей, $\epsilon=10^{-3}$.

\begin{table}[h]
	\caption{Результаты решения серии задач с разрывными функциями}
	\begin{center}
		\begin{tabular}{|c|c|c|}
			\hline
			Метод & Решено задач & Среднее число испытаний \\
			\hline
			\hline
			ИО    &  89 &   876  \\
			\hline
			АГП-Р & 100  & 81 \\
			\hline
		\end{tabular}
	\end{center}
\end{table}

Результаты показывают, что метод АГП-Р успешно решил все задачи серии, тогда как метод имитации отжига не справился с решением 11 задач, затратив при этом в среднем в 10 раз больше испытаний.

\begin{biblio}

%\bibitem{fio_bib1} Шаманаев П. А., Язовцева О. С. Достаточные условия локальной покомпонентной асимптотической эквивалентности нелинейных систем обыкновенных дифференциальных уравнений и ее приложение к устойчивости по части переменных // Журнал Средневолжского математического общества. 2017. Т. 19, № 1. С. 102--115.

%\bibitem{fio_bib2} Шаманаев П. А., Язовцева О. С. Достаточные условия полиустойчивости по части переменных нулевого решения нелинейных систем обыкновенных дифференциальных уравнений // Журнал Средневолжского математического общества.  2018. Т. 20, № 3. С. 304--317.

%\bibitem{fio_bib3} Шаманаев П. А., Язовцева О. С. Исследование устойчивости положения равновесия системы динамики биоценоза в условиях межвидового взаимодействия // Вестник Мордовского университета. 2018. Т. 28, № 3. С. 321–332.

%\bibitem{fio_bib4} Малкин И. Г. Теория устойчивости движения. М.: Наука, 1966. 533 с.

\bibitem{Grishagin2007} 
Городецкий~С.Ю., Гришагин~В.А. Нелинейное программирование и многоэкстремальная оптимизация. Н.~Новгород: Изд-во ННГУ, 2007. 489~с.

\bibitem{Sergeyev2008}
Сергеев Я.Д., Квасов Д.Е. Диагональные методы глобальной оптимизации. М.: Физматлит, 2008. 352 с.

\bibitem{Evtushenko2009}
Евтушенко Ю.Г., Малкова В.У., Станевичюс А.-И.А. Параллельный поиск глобального экстремума функций многих переменных // Ж. вычисл. матем. и матем. физ. 2009. Т. 49, № 2.  С. 255--269.

\bibitem{Elsakov}
Елсаков С.М., Ширяев В.И. Однородные алгоритмы многоэкстремальной оптимизации // Ж. вычисл. матем. и матем. физ. 2010. Т. 50, № 10. С. 1727--1740.

\bibitem{Batuhtin1997}
Батухтин В.Д., Бигильдеев С.И., Бигильдеева Т.Б. Численные методы решения разрывных экстремальных задач // Изв. РАН, сер. теория и сист. управления. 1997. Т. 3. С. 113--120.

\bibitem{Jihui}
Jihui Z., Junqin X. A new differential evolution for discontinuous optimization problems // 
Proceedings of the Third International Conference on Natural Computation. 2007. Vol. 3. P. 483--487.

\bibitem{Strongin2013}
Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А. Параллельные вычисления в задачах глобальной оптимизации. М.: Изд-во МГУ, 2013. 280 с.

\end{biblio}

\msc{90C26}

\title{Solving global optimization problems taking into account possible discontinuities of the objective function}

\authors{K.A. Barkalov, M.A. Usova}
\organizations{ Lobachevsky State University of Nizhni Novgorod }

\end{document}

### 2020-06-26 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
fix some obvious typos

@@
 This approach is also widely known as the $\varepsilon$-constraint method (or the method of successive concessions) \cite{c39,c40,c41,c42}.
-The choice of feasible concessions $\delta_i$, $1 \leq i \leq s$ in the computational scheme (\ref{eq:9}) allows obtaining any efficient decision $\widetilde{y}^* \in P_{lex}(f,D)$ and takes into account the peculiarities of the MCOlex problem.
+The choice of feasible concessions $\delta_i$, $1 \leq i \leq s$, in the computational scheme (\ref{eq:9}) allows obtaining any efficient decision $\widetilde{y}^* \in P_{lex}(f,D)$ and takes into account the peculiarities of the MCOlex problem.
 At the same time, this approach makes it necessary, at each stage of the scheme (\ref{eq:9}), to solve more complex global optimization problems with nonlinear constraints.
@@
 
-The first trial is carried out at an arbitrary internal point $x^1\in(0,1)$. The choice of the point $x^{k+1}$, $k \geq 1$ of any next trial is determined by the following rules
+The first trial is carried out at an arbitrary internal point $x^1\in(0,1)$. The choice of the point $x^{k+1}$, $k \geq 1$, of any next trial is determined by the following rules.
 
@@
 \begin{equation}\label{eq:27}
-x^{k+1} = \mathcal{X}(x_{t-1},x_t)
+x^{k+1} = \mathcal{X}(x_{t-1},x_t),
 \end{equation}
@@
 Iterations of the algorithm are terminated if the stop criterion is satisfied
-
 \begin{equation}\label{eq:28}
@@
 \begin{enumerate}
-	\item the functions $g_j$, $1 \leq j \leq m+1$, $[0,1]$ satisfy the H\"older condition of (\ref{eq:13}) with the constants $H_j$, $1 \leq j \leq m+1$,
-	\item for the values $H^k_\nu$ from (\ref{eq:24}), starting from a certain step, the inequalities are valid\footnote{The values $r_{\nu} > 1$, $1 \leq \nu \leq m+1$,are the AGCS reliability parameters used for computing the interval characteristics in (\ref{eq:25}).}
+	\item the functions $g_j(x)$, $1 \leq j \leq m+1$, $x\in[0,1]$, satisfy the H\"older condition (\ref{eq:13}) with the constants $H_j$, $1 \leq j \leq m+1$,
+	\item for the values $H^k_\nu$ from (\ref{eq:24}), starting from a certain step, the inequalities are valid\footnote{The values $r_{\nu} > 1$, $1 \leq \nu \leq m+1$, are the AGCS reliability parameters used for computing the interval characteristics in (\ref{eq:25}).}
 \begin{equation} \label{eq:29}
@@
 With a complete scheme for calculating criteria values, i.e.
-when at the iteration points $x_i$, $1 \leq i \leq k$ the values of all criteria $f_j(x_i)$, $1 \leq j \leq s$, $1 \leq i \leq k$, are calculated, the effect of using the search information $A_k$ can be even more significant.
-In this case, the search information can also be used when changing the values of the concessions $\delta_j$, $1 \leq j < s$ and thus, the solution of the next problem from the $\mathbb{P}_t$ family (\ref{eq:10}) can be carried out repeatedly using the results of all previously performed calculations.
+when at the iteration points $x_i$, $1 \leq i \leq k$, the values of all criteria $f_j(x_i)$, $1 \leq j \leq s$, $1 \leq i \leq k$, are calculated, the effect of using the search information $A_k$ can be even more significant.
+In this case, the search information can also be used when changing the values of the concessions $\delta_j$, $1 \leq j < s$, and thus, the solution of the next problem from the $\mathbb{P}_t$ family (\ref{eq:10}) can be carried out repeatedly using the results of all previously performed calculations.
 The results of numerical experiments (see Section \ref{sec:5}), demonstrate that the volume of computations performed, based on reusing search information, can significantly decrease.
@@
 updated February 2010 (2009)
-\bibitem{c55} Cust\'odio, A.L., Aguilar, J.F. Madeira, Ismael Vaz, Lu\'ls N. Vicente: Direct Multisearch for Multiobjective Optimization. IAM Journal on Optimization 21(3). pp. 1109-1140 (2011)
+\bibitem{c55} Cust\'odio, A.L., Aguilar, J.F. Madeira, Ismael Vaz, Lu\'ls N. Vicente: Direct Multisearch for Multiobjective Optimization. SIAM J. Optim. 21(3). pp. 1109--1140 (2011)
 \bibitem{c56} Brockhoff, D., Tusar, T., Auger, A., Hansen, N.: Using Well-Understood Single-Objective Functions in Multiobjective Black-Box Optimization Test Suites.
@@
 298--311 (2018)
-\bibitem{c59} Lai, L.
-, Fiaschi, L.
-, Cococcioni, M.: Solving Mixed Pareto-Lexicographic Multi-Objective Optimization Problems: The Case of Priority Chains, Swarm and Evolutionary Computation, 55.
+\bibitem{c59} Lai, L., Fiaschi, L., Cococcioni, M.: Solving Mixed Pareto-Lexicographic Multi-Objective Optimization Problems: The Case of Priority Chains, Swarm and Evolutionary Computation, 55.
 pp.

### 2020-06-27 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
some minor corrections

@@
 \email{konstantin.barkalov@itmm.unn.ru} \\
-\at Institute of Informational Technologies, Mathematics and Mechanics, Mathematical Center, Lobachevsky State University of Nizhni Novgorod, Gagarin ave., 23, Nizhni Novgorod, Russia 
+\at Mathematical Center, Lobachevsky State University of Nizhni Novgorod, Gagarin ave., 23, Nizhni Novgorod, Russia 
 }
@@
 The results of research have been published in a large number of monographs -- see, for example, \cite{c17,c18,c19,c20,c21,c22,c23,c24,c25}.
-Making a reasonable choice of the approach for solving GO problems from the entire set of already existing global search algorithms requires taking into account the following:
-
+Making a reasonable choice of the approach for solving GO problems from the entire set of already existing global search algorithms requires taking into account the following.
 \begin{itemize}
-	\item Since the estimate of the global minimum is an integral characteristic of the optimization problem (the value of the minimized function at the point of the global minimum should be compared with the values of the function at all other points of the search domain), the global search methods should provide at least some degree of coverage of the search domain,
+	\item Since the estimate of the global minimum is an integral characteristic of the optimization problem (the value of the minimized function at the point of the global minimum should be compared with the values of the function at all other points of the search domain), the global search methods should provide at least some degree of coverage of the search domain.
 	\item Due to the ``curse of dimensionality'' reducing computational complexity requires the resulting coverage of the search domain to be heterogeneous (nonuniform), i.e.
-it must be dense only in the vicinity of globally optimal points,
-	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- on one of the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}),
+it must be dense only in the vicinity of globally optimal points.
+	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- one of the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}).
 	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function at these points), which significantly increases the complexity and computational cost of the global optimization methods.
-Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47}, the Peano space-filling curves \cite{c17,c23},
+Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47}, the Peano space-filling curves \cite{c17,c23}.
 	\item When solving multiple problems of multistage lexicographic multicriteria optimization, global search methods can reduce the complexity of calculations by using search information obtained during the calculations.
@@
 Given the requirements listed above, the proposed approach for solving MMCOlex problems is based on three key ideas:
-
 \begin{itemize}
@@
 
-The presence of nonlinear constraints significantly increasesthe complexity of global optimization problems; the resulting solutions must belong to a feasible domain
+The presence of nonlinear constraints significantly increases the complexity of global optimization problems; the resulting solutions must belong to a feasible domain
 \begin{equation}\label{eq:17}
@@
 \begin{enumerate}
-	\item the functions $g_j(x)$, $1 \leq j \leq m+1$, $x\in[0,1]$, satisfy the H\"older condition (\ref{eq:13}) with the constants $H_j$, $1 \leq j \leq m+1$,
+	\item the functions $g_j(x)$, $1 \leq j \leq m+1$, $x\in[0,1]$, satisfy the H\"older condition with the constants $H_j$, $1 \leq j \leq m+1$,
 	\item for the values $H^k_\nu$ from (\ref{eq:24}), starting from a certain step, the inequalities are valid\footnote{The values $r_{\nu} > 1$, $1 \leq \nu \leq m+1$, are the AGCS reliability parameters used for computing the interval characteristics in (\ref{eq:25}).}
@@
 \begin{theorem}
-Let the next solved optimization problem $\psi(x)$ of the computational scheme (\ref{eq:9}) differ from the previously solved problem $\phi(x)$ by some limited value $\delta$, i.e.
+Let the next minimized function $\psi(x)$ of the computational scheme (\ref{eq:9}) differ from the previous minimized function $\phi(x)$ by some limited value $\delta$, i.e.
 \begin{equation}\label{eq:32}
-|\psi(x)-\phi(x)| \leq \delta, x \in [0,1]
+|\psi(x)-\phi(x)| \leq \delta, x \in [0,1],
 \end{equation}
@@
 \begin{equation}\label{eq:38}
-x_k^* = arg \min{\phi(x_k)}, 1 \leq i \leq k, 
+x_k^* = arg \min{\phi(x_k)}, 1 \leq i \leq k.
 \end{equation}
-\end{theorem}(the points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $H_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively).
+\end{theorem}
+The points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $H_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively.
 
@@
 \begin{theorem}
-Let the optimization problem $\phi(x)$ of the computational scheme (\ref{eq:9}) be solved before the stopping criterion in (\ref{eq:28}) is fulfilled for the specified accuracy of the global search $\varepsilon > 0$ and the density of trials for the interval $[a,b] \subset [0,1]$ is reached
+Let the function $\phi(x)$ of the computational scheme (\ref{eq:9}) be minimized before the stopping criterion in (\ref{eq:28}) is fulfilled for the specified accuracy of the global search $\varepsilon > 0$ and the density of trials $p_{ab}$ for the interval $[a,b] \subset [0,1]$ is reached
 \begin{equation}\label{eq:39}
-p_{ab}=\frac{n_{ab}}{b-a}
+p_{ab}=\frac{n_{ab}}{b-a},
 \end{equation}
 where $n_{ab}$ is the number of trial points in the interval $[a,b]$ during the minimization of the function $\phi(x)$.
-Then, when solving the next optimization problem $\psi(x)$ using the search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, the number of trials in the same interval $[a,b]$ instead of the maximum value
+Then, when minimizing the next function $\psi(x)$ using the search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, the number of trials in the same interval $[a,b]$ instead of the maximum value
 \begin{equation}\label{eq:40}
-(3m/2\Delta)(b-a), m \; \text{from} \;  (\ref{eq:35}),
+(3m/2\Delta)(b-a), \text{where} \; m \; \text{is from} \;  (\ref{eq:35}),
 \end{equation}
@@
 \begin{theorem}
-If the number of trials in the interval $[a,b]\subset[0,1]$ exceeds the maximum estimate (\ref{eq:41}) when solving the optimization problem $\psi(x)$ of the computational scheme (\ref{eq:9}), then when solving the next optimization problem $\psi(x)$ using the search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, additional trials in the interval $[a,b]$ will not be performed while the condition (\ref{eq:42}) is fulfilled.
+If the number of trials in the interval $[a,b]\subset[0,1]$ exceeds the maximum estimate (\ref{eq:41}) when minimizing the function $\phi(x)$ of the computational scheme (\ref{eq:9}), then when minimizing the next function $\psi(x)$ using the search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, additional trials in the interval $[a,b]$ will not be performed while the condition (\ref{eq:42}) is fulfilled.
 \end{theorem}
@@
 
-It should be noted that the number of iterations of the global search is also reduced in the course of sequential solution of subproblems of the same MCO problems.
+It should be noted that the number of iterations of the global search is also reduced in the course of sequential solution of subproblems of the same MCO problem.
 

### 2020-06-28 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
OptLet правка списка литературы

@@
 \bibitem{c5} Pardalos, P.M., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-Convex Multi-Objective Optimization. Springer. (2017)
-\bibitem{c6} Marler, R. T., Arora, J. S.: Survey of multi-objective optimization methods for engineering. Struct. Multidisciplinary Optimization 26, pp. 369--395 (2004)
+\bibitem{c6} Marler, R., Arora , J.: Survey of multi-objective optimization methods for engineering. Struct. Multidisc. Optim. \textbf{26}, 369--395 (2004)
 \bibitem{c7} Figueira,J., Greco, S., Ehrgott, M., editors.: Multiple criteria decision analysis: State of the art surveys. Springer, New York (NY). (2005)
-\bibitem{c8} Zavadskas, E. K., Turskis, Z., Kildien\.e, S.: State of art surveys of overviews on MCDM/MADM methods. Technological and Economic Development of Economy, 20, pp. 165--179 (2014)
-\bibitem{c9} Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications. Surv. Math. Ind. 11, pp. 1--42 (2005)
-\bibitem{c10} Branke, J., Deb, K., Miettinen, K., Slowinski, R., editors.: Multi-Objective Optimization—Interactive and Evolutionary Approaches. Springer, Berlin. (2008)
+\bibitem{c8} Zavadskas, E.K., Turskis, Z., Kildien\.e, S.: State of art surveys of overviews on MCDM/MADM methods. Technol. Econ. Dev. Econ. \textbf{20}, 165--179 (2014)
+\bibitem{c9} Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications. Surv. Math. Ind. \textbf{11}, 1--42 (2005)
+\bibitem{c10} Branke, J., Deb, K., Miettinen, K., Slowinski, R., editors.: Multi-Objective Optimization--Interactive and Evolutionary Approaches. Springer, Berlin. (2008)
 \bibitem{c11} Deb, K.: Multi-Objective Optimization using Evolutionary Algorithms. Wiley, Chichester. (2001)
-\bibitem{c12} Yang, X.-S.: Nature-inspired metaheuristic algorithms. Luniver Press, Frome. (2008) 
+\bibitem{c12} Yang, X.-S.: Nature-inspired Metaheuristic Algorithms. Luniver Press, Frome. (2008) 
 \bibitem{c13} Tan, K.C., Khor, E.F., Lee, T.H.: Multi-objective Evolutionary Algorithms and Applications. Springer-Verlag, London. (2005)
-\bibitem{c14} Eichfelder, G.: Scalarizations for adaptively solving multi-objective optimization problems. Comput. Optim. Appl. 44, pp. 249--273 (2009)
-\bibitem{c15} Jones, D.R.: A taxonomy of global optimization methods based on response surfaces. J. Glob. Optim. 21, pp. 345--383 (2001)
-\bibitem{c16} Voutchkov, I., Keane, A.: Multi-objective optimization using surrogates. Comput. Intell. Optim. Adapt. Learn. Optim. 7, pp. 155--175 (2010)
-\bibitem{c17} Strongin, R.G.: Numerical methods in multiextremal problems: information-statistical algorithms. Nauka, Moscow. (1978, in Russian) 
+\bibitem{c14} Eichfelder, G.: Scalarizations for adaptively solving multi-objective optimization problems. Comput. Optim. Appl. \textbf{44}, 249--273 (2009)
+\bibitem{c15} Jones, D.R.: A taxonomy of global optimization methods based on response surfaces. J. Glob. Optim. \textbf{21}, 345--383 (2001)
+\bibitem{c16} Voutchkov, I., Keane, A.: Multi-objective optimization using surrogates. Comput. Intell. Optim. Adapt. Learn. Optim. \textbf{7}, 155--175 (2010)
+\bibitem{c17} Strongin, R.G.: Numerical Methods in Multiextremal Problems: Information-statistical Algorithms. Nauka, Moscow. (1978, in Russian) 
 \bibitem{c18} Strongin, R., Sergeyev, Ya.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht. (2nd ed. 2013, 3rd ed. 2014).
@@
 \bibitem{c21} Zhigljavsky, A.A.: Theory of Global Random Search. Kluwer Academic Publishers, Dordrecht. (1991)
-\bibitem{c22} Pint\'er, J.D.: Global optimization in Action (continuous and Lipschitz optimization: algorithms, implementations and applications). Kluwer Academic Publishers, Dortrecht. (1996)
-\bibitem{c23} Sergeyev Y.D., Strongin R.G., Lera D.: Introduction to global optimization exploiting space-filling curves. Springer. (2013)
+\bibitem{c22} Pint\'er, J.D.: Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications). Kluwer Academic Publishers, Dortrecht. (1996)
+\bibitem{c23} Sergeyev, Y.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting Space-filling Curves. Springer. (2013)
 \bibitem{c24} Locatelli, M., Schoen, F.: Global Optimization: Theory, Algorithms, and Applications. SIAM. (2013)
 \bibitem{c25} Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. Princeton University Press. (2016)
-\bibitem{c26} Lera D., Sergeyev Ya.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and Holder constants, Communications in Nonlinear Science and Numerical Simulation, 23.
-pp.
-328--342 (2015)
-\bibitem{c27} Gergel, V.: A Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems.
-2nd International Conference on Mathematics and Computers in Sciences and in Industry.
-pp.
-13--18 (2015).
-DOI: 10.1109/MCSI.2015.18
+\bibitem{c26} Lera D., Sergeyev Ya.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and Holder constants.
+Commun.
+Nonlinear Sci.
+\textbf{23}(1-3), 328--342 (2015)
+\bibitem{c27} Gergel, V.: A unified approach to use of coprocessors of various types for solving global optimization problems. 2nd Int. Conf. on MACSI. 13--18 (2015). DOI: 10.1109/MCSI.2015.18
 \bibitem{c28} Arora, R.K.: Optimization: Algorithms and Applications. CRC Press. (2015)
-\bibitem{c29} Bazaraa, M.S., Sherali, H.D., Shetty, C. M.: Nonlinear Programming: Theory and Algorithms, John Wiley and Sons. (2006)
+\bibitem{c29} Bazaraa, M.S., Sherali, H.D., Shetty, C.M.: Nonlinear Programming: Theory and Algorithms. John Wiley and Sons. (2006)
 \bibitem{c30} Gergel, V.P., Kozinov, E.A.: Accelerating multicriterial optimization by the intensive exploitation of accumulated search data.
-In: AIP Conference Proceedings, 1776.
-pp.
-090003 (2016).
+In: 	AIP Conf.
+Proc., \textbf{1776}, 090003 (2016).
 DOI: 10.1063/1.4965367
@@
 Glob.
-Optim., 71(1).
-pp.
-73--90 (2018).
+Optim.
+\textbf{71}(1), 73--90 (2018).
 DOI: 10.1007/s10898-018-0624-3
 \bibitem{c32} Sysoyev, A., Barkalov, K., Gergel, V.: Globalizer: A novel supercomputer software system for solving time-consuming global optimization problems.
-Numerical Algebra, Control and Optimization 8(1), pp.
-47--62 (2018)
-\bibitem{c33} Kvasov D.E., Sergeyev Y.D.: Deterministic approaches for solving practical black-box global optimization problems. Advances in Engineering Software, 80, pp.58--66 (2015)
-\bibitem{c34} Modorskii, V., Gaynutdinova, D., Gergel, V., Barkalov, K.: Optimization in design of scientfic products for purposes of cavitation problems. AIP Conference Proceedings, 1738. (2016)
-\bibitem{c35} Evtushenko, Yu.G., Posypkin, M.A.: A deterministic algorithm for global multi-objective optimization. Optimization Methods \& Software, 29(5), pp. 1005--1019 (2014)
+Numer.
+Algebra, Control.
+Optim.
+\textbf{8}(1), 47--62 (2018)
+\bibitem{c33} Kvasov D.E., Sergeyev Y.D.: Deterministic approaches for solving practical black-box global optimization problems. Adv. Eng. Softw. \textbf{80}, 58--66 (2015)
+\bibitem{c34} Modorskii, V., Gaynutdinova, D., Gergel, V., Barkalov, K.: Optimization in design of scientfic products for purposes of cavitation problems. AIP Conf. Proc. \textbf{1738}, 400013 (2016)
+\bibitem{c35} Evtushenko, Yu.G., Posypkin, M.A.: A deterministic algorithm for global multi-objective optimization. Optim. Methods Softw., \textbf{29}(5), 1005--1019 (2014)
 \bibitem{c36} {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Adaptation of a one-step worst-case optimal univariate algorithm of bi-objective Lipschitz optimization to multidimensional problems.
-Commun Nonlinear Sci Numer Simulat, 21, pp.
-89--98 (2015)
+Commun.
+Nonlinear Sci.
+Numer.
+Simul., \textbf{21}(1-3), 89--98 (2015)
 \bibitem{c37} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
-ACM Transactions on Mathematical Software 29(4).
-pp.
-469--480 (2003)
+ACM Trans.
+Math.
+Softw.
+\textbf{29}(4), 469--480 (2003)
 
-\bibitem{c38} Zarepisheh, M., Khorram, E.: On the transformation of lexicographic nonlinear multiobjective  programs to single objective programs. Math. Methods Oper. Res. 74. pp. 217--231 (2011)
+\bibitem{c38} Zarepisheh, M., Khorram, E.: On the transformation of lexicographic nonlinear multiobjective  programs to single objective programs. Math. Methods Oper. Res. \textbf{74}, 217--231 (2011)
 \bibitem{c39} Haimes, Y.Y., Lasdon, L., Wismer, D.: On a bicriterion formulation of the problem of integrated systems identification and system optimization.
@@
 Man Cybern.
-1.
-pp.
-296--297 (1971)
-\bibitem{c40} Chankong, V., Haimes, Y.Y.: Multiobjective decision making: theory and methodology. North-Holland Series in System Science and Engineering. New York (NY): Elsevier. (1983)
-\bibitem{c41} Ehrgott, M., Ruzika, S.: Improved $\varepsilon$-constraint method for multiobjective programming. J. Optim. Theory Appl. 138. pp. 375--396 (2008)
+\textbf{1}, 296--297 (1971)
+\bibitem{c40} Chankong, V., Haimes, Y.Y.: Multiobjective Decision Making: Theory and Methodology. North-Holland Series in System Science and Engineering. New York (NY): Elsevier. (1983)
+\bibitem{c41} Ehrgott, M., Ruzika, S.: Improved $\varepsilon$-constraint method for multiobjective programming. J. Optim. Theory Appl. \textbf{138}, 375--396 (2008)
 \bibitem{c42} Rastegar, N., Khorram, E.: Relaxation of constraints in lexicographic multiobjective programming problems.
-A Journal of Mathematical Programming and Operations Research, 64.
-pp.
-2111--2129 (2015)
-\bibitem{c43} Castro-Gutierrez, J., Landa-Silva, D., Pérez, J.M.: Improved Dynamic Lexicographic Ordering for Multi-Objective Optimisation.
-In: Schaefer R., Cotta C., Kolodziej J., Rudolph G.
-(eds) Parallel Problem Solving from Nature, PPSN XI.
-PPSN 2010.
-Lecture Notes in Computer Science, 6239.
-pp 31-40 (2010)
-\bibitem{c44} Audet, C., Savard, G., Zghal, W.: Multiobjective Optimization Through a Series of Single-Objective Formulations SIAM J. Optim.,19(1). pp. 188--210 (2008)
-\bibitem{c45} Sergeyev, Y.D., Kvasov, D.E.: Deterministic Global Optimization: An Introduction to the Diagonal Approach (SpringerBriefs in Optimization). Springer. (2017)
+A Journal of Mathematical Programming and Operations Research, \textbf{64}, 2111--2129 (2015)
+\bibitem{c43} Castro-Gutierrez, J., Landa-Silva, D., P\'erez, J.M.: Improved dynamic lexicographic ordering for multi-objective optimisation. Lect. Notes Comput. Sci., \textbf{6239}, 31--40 (2010) 
+\bibitem{c44} Audet, C., Savard, G., Zghal, W.: Multiobjective optimization through a series of single-objective formulations. SIAM J. Optim. \textbf{19}(1), 188--210 (2008)
+\bibitem{c45} Sergeyev, Y.D., Kvasov, D.E.: Deterministic Global Optimization: An Introduction to the Diagonal Approach (Springer Briefs in Optimization). Springer. (2017)
 \bibitem{c46} Carr, C.R., Howe, C.W.: Quantitative Decision Procedures in Management and Economic: Deterministic Theory and Applications. New York: McGraw–Hill (1964)
-\bibitem{c47} Dam, E.R., Husslage, B., Hertog, D.: One-dimensional Nested Maximin Designs. Journal of Global Optimization, 46(2). pp. 287--306 (2010)
-\bibitem{c48} Gergel, V., Kozinov, E.: Efficient multicriterial optimization based on intensive reuse of search information, J. Glob. Optim., 71(1). pp. 73--90 (2018). doi:10.1007/s10898-018-0624-3.
-\bibitem{c49} Mart\'l, L., Garc\'la, J., Berlanga, A., Molina, J.M.: A stopping criterion for multi-objective optimization evolutionary algorithms, Information Sciences 367, pp. 700--718 (2016)
-\bibitem{c50} Dilettoso, E., Rizzo, S.A., Salerno, N.: A weakly pareto compliant quality indicator. Mathematical and Computational Applications 22(1), 25 (2017)
-\bibitem{c51} Wu, J., Azarm, S.: Metrics for quality assessment of a multiobjective design optimization solution set, Journal of Mechanical Design 123(1). pp. 18--25 (2001)
-\bibitem{c52} Audet, C., Bigeon, J., Cartier, D., Digabel,  S. Le, Salomon, L.: Performance indicators in multiobjective optimization. G-2018-90, (2018)
-\bibitem{c53} Huband, S., Hingston, P., Barone, L., and While, L.: A Review of Multiobjective Test Problems and a Scalable Test Problem Toolkit.
-IEEE Transactionson Evolutionary Computation, 10(5).
-pp.
-477--506 (2006)
+\bibitem{c47} Dam, E.R., Husslage, B., Hertog, D.: One-dimensional nested maximin designs. J. Glob. Optim. \textbf{46}(2), 287--306 (2010)
+\bibitem{c48} Gergel, V., Kozinov, E.: Efficient multicriterial optimization based on intensive reuse of search information, J.
+Glob.
+Optim., \textbf{71}(1), 73--90 (2018).
+doi:10.1007/s10898-018-0624-3.
+\bibitem{c49} Mart\'l, L., Garc\'la, J., Berlanga, A., Molina, J.M.: A stopping criterion for multi-objective optimization evolutionary algorithms. Inf. Sci. \textbf{367--368}, 700--718 (2016)
+\bibitem{c50} Dilettoso, E., Rizzo, S.A., Salerno, N.: A weakly pareto compliant quality indicator. Math. Comput. Appl. \textbf{22}(1), 25 (2017)
+\bibitem{c51} Wu, J., Azarm, S.: Metrics for quality assessment of a multiobjective design optimization solution set. J. Mech. Des. \textbf{123}(1), 18--25 (2001)
+\bibitem{c52} Audet, C., Bigeon, J., Cartier, D., Digabel, S.Le, Salomon, L.: Performance indicators in multiobjective optimization. G-2018-90, (2018)
+\bibitem{c53} Huband, S., Hingston, P., Barone, L., While, L.: A review of multiobjective test problems and a scalable test problem toolkit. IEEE T. Evolut. Comput. \textbf{10}(5), 477--506 (2006)
 \bibitem{c54} Hansen, N., Auger, A., Finck, S., and Ros, R.: Real-Parameter Black-Box Optimization Benchmarking 2009: Experimental Setup.
-INRIA Research Report RR-6829, INRIA Saclay—Ile-de-France.
+INRIA Research Report RR-6829, INRIA Saclay-Ile-de-France.
 updated February 2010 (2009)
-\bibitem{c55} Cust\'odio, A.L., Aguilar, J.F. Madeira, Ismael Vaz, Lu\'ls N. Vicente: Direct Multisearch for Multiobjective Optimization. SIAM J. Optim. 21(3). pp. 1109--1140 (2011)
-\bibitem{c56} Brockhoff, D., Tusar, T., Auger, A., Hansen, N.: Using Well-Understood Single-Objective Functions in Multiobjective Black-Box Optimization Test Suites.
+\bibitem{c55} Cust\'odio, A.L., Madeira, J.F.A., Vaz, A.I.F., Vicente, L.N.: Direct multisearch for multiobjective optimization. SIAM J. Optim. \textbf{21}(3), 1109--1140 (2011)
+\bibitem{c56} Brockhoff, D., Tusar, T., Auger, A., Hansen, N.: Using well-understood single-objective functions in multiobjective black-box optimization test suites.
 (2019) https://arxiv.org/abs/1604.00359
-\bibitem{c57} Cococcioni, M., Cudazzo, A., Pappalardo, M., Sergeyev, Y.D.: Solving the Lexicographic Multi-Objective Mixed-Integer Linear Programming Problem using branch-and-bound and grossone methodology, Communications in Nonlinear Science and Numerical Simulation, 84.
-pp.
-105177 (2020)
-\bibitem{c58} Cococcioni, M., Pappalardo, M., Sergeyev, Y.D.: Lexicographic multi-objective linear programming using grossone methodology: Theory and algorithm, Applied Mathematics and Computation, 318.
-pp.
-298--311 (2018)
-\bibitem{c59} Lai, L., Fiaschi, L., Cococcioni, M.: Solving Mixed Pareto-Lexicographic Multi-Objective Optimization Problems: The Case of Priority Chains, Swarm and Evolutionary Computation, 55.
-pp.
-100687 (2020) DOI: 10.1016/j.swevo.2020.100687
+\bibitem{c57} Cococcioni, M., Cudazzo, A., Pappalardo, M., Sergeyev, Y.D.: Solving the lexicographic multi-objective mixed-integer linear programming problem using branch-and-bound and grossone methodology.
+Commun.
+Nonlinear Sci.
+Numer.
+Simul.
+\textbf{84}, 105177 (2020)
+\bibitem{c58} Cococcioni, M., Pappalardo, M., Sergeyev, Y.D.: Lexicographic multi-objective linear programming using grossone methodology: Theory and algorithm.
+Appl.
+Math.
+Comput.
+\textbf{318}, 298--311 (2018)
+\bibitem{c59} Lai, L., Fiaschi, L., Cococcioni, M.: Solving mixed pareto-lexicographic multi-objective optimization problems: the case of priority chains.
+Swarm Evol.
+Comput.
+\textbf{55}, 100687 (2020) DOI: 10.1016/j.swevo.2020.100687
 

### 2020-06-29 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
OptLet правки списка литературы

@@
 \bibitem{c6} Marler, R., Arora , J.: Survey of multi-objective optimization methods for engineering. Struct. Multidisc. Optim. \textbf{26}, 369--395 (2004)
-\bibitem{c7} Figueira,J., Greco, S., Ehrgott, M., editors.: Multiple criteria decision analysis: State of the art surveys. Springer, New York (NY). (2005)
+\bibitem{c7} Figueira,J., Greco, S., Ehrgott, M., (eds.): Multiple criteria decision analysis: State of the art surveys. Springer, New York (NY). (2005)
 \bibitem{c8} Zavadskas, E.K., Turskis, Z., Kildien\.e, S.: State of art surveys of overviews on MCDM/MADM methods. Technol. Econ. Dev. Econ. \textbf{20}, 165--179 (2014)
 \bibitem{c9} Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications. Surv. Math. Ind. \textbf{11}, 1--42 (2005)
-\bibitem{c10} Branke, J., Deb, K., Miettinen, K., Slowinski, R., editors.: Multi-Objective Optimization--Interactive and Evolutionary Approaches. Springer, Berlin. (2008)
+\bibitem{c10} Branke, J., Deb, K., Miettinen, K., Slowinski, R., (eds.): Multi-Objective Optimization--Interactive and Evolutionary Approaches. Springer, Berlin. (2008)
 \bibitem{c11} Deb, K.: Multi-Objective Optimization using Evolutionary Algorithms. Wiley, Chichester. (2001)
@@
 \bibitem{c17} Strongin, R.G.: Numerical Methods in Multiextremal Problems: Information-statistical Algorithms. Nauka, Moscow. (1978, in Russian) 
-\bibitem{c18} Strongin, R., Sergeyev, Ya.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht. (2nd ed. 2013, 3rd ed. 2014).
+\bibitem{c18} Strongin, R.G., Sergeyev, Y.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht. (2nd ed. 2013, 3rd ed. 2014).
 \bibitem{c19} T\"orn, A., {\v Z}ilinskas, A.: Global Optimization. Lecture Notes in Computer Science, 350. Springer-Verlag, Berlin. (1989)
@@
 \bibitem{c25} Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. Princeton University Press. (2016)
-\bibitem{c26} Lera D., Sergeyev Ya.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and Holder constants.
+\bibitem{c26} Lera, D., Sergeyev, Y.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and Holder constants.
 Commun.
@@
 \textbf{8}(1), 47--62 (2018)
-\bibitem{c33} Kvasov D.E., Sergeyev Y.D.: Deterministic approaches for solving practical black-box global optimization problems. Adv. Eng. Softw. \textbf{80}, 58--66 (2015)
+\bibitem{c33} Kvasov, D.E., Sergeyev, Y.D.: Deterministic approaches for solving practical black-box global optimization problems. Adv. Eng. Softw. \textbf{80}, 58--66 (2015)
 \bibitem{c34} Modorskii, V., Gaynutdinova, D., Gergel, V., Barkalov, K.: Optimization in design of scientfic products for purposes of cavitation problems. AIP Conf. Proc. \textbf{1738}, 400013 (2016)
@@
 Simul., \textbf{21}(1-3), 89--98 (2015)
-\bibitem{c37} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
+\bibitem{c37} Gaviano, M., Kvasov, D.E, Lera, D., Sergeyev, Y.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
 ACM Trans.
@@
 \bibitem{c51} Wu, J., Azarm, S.: Metrics for quality assessment of a multiobjective design optimization solution set. J. Mech. Des. \textbf{123}(1), 18--25 (2001)
-\bibitem{c52} Audet, C., Bigeon, J., Cartier, D., Digabel, S.Le, Salomon, L.: Performance indicators in multiobjective optimization. G-2018-90, (2018)
+\bibitem{c52} Audet, C., Bigeon, J., Cartier, D., Digabel, S.Le, Salomon, L.: Performance indicators in multiobjective optimization. Technical Report G-2018-90, Les cahiers du GERAD. (2018)
 \bibitem{c53} Huband, S., Hingston, P., Barone, L., While, L.: A review of multiobjective test problems and a scalable test problem toolkit. IEEE T. Evolut. Comput. \textbf{10}(5), 477--506 (2006)
-\bibitem{c54} Hansen, N., Auger, A., Finck, S., and Ros, R.: Real-Parameter Black-Box Optimization Benchmarking 2009: Experimental Setup.
+\bibitem{c54} Hansen, N., Auger, A., Finck, S., Ros, R.: Real-Parameter Black-Box Optimization Benchmarking 2009: Experimental Setup.
 INRIA Research Report RR-6829, INRIA Saclay-Ile-de-France.

### 2020-06-29 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
one correction

@@
 \begin{theorem}
-Let the next minimized function $\psi(x)$ of the computational scheme (\ref{eq:9}) differ from the previous minimized function $\phi(x)$ by some limited value $\delta$, i.e.
+Let the next minimized function $\psi(x)$ of the computational scheme (\ref{eq:9}) differs from the previous minimized function $\phi(x)$ by some limited value $\delta$, i.e.
 \begin{equation}\label{eq:32}

### 2020-07-02 2020/Saransk_Yamshchikov/theses.tex
Saransk thesis No 2

% !!!!!!!!!!!!!!!!!!   ВНИМАНИЕ !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
% Заголовки разделов формируются при помощи команд \section{}, \subsection{}, \subsubsection{}
% Не используйте уровень вложенности заголовков больше трех!
% -----------------------------
% Для оформления теорем, лемм, следствий используйте окружения 
% Def     - Определение
% Teor    - Теорема
% Lem     - Лемма
% Predl   - Предложение
% Ass     - Утверждение
% Cor     - Следствие
% Example - Пример
% -----------------------------
% Доказательство теоремы начинается командой \proof и завершается командой \endproof
% -----------------------------
% Литература помещается в окружение biblio.

\documentclass[11pt, oneside, a4paper]{article}
%\usepackage[cp1251]{inputenc} % кодировка
\usepackage[utf8]{inputenc} % кодировка
\usepackage[english, russian]{babel} % Русские и английские переносы
\usepackage{graphicx}          % для включения графических изображений
\usepackage{cite}              % для корректного оформления литературы
\usepackage{enumitem}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{mathtext}

%стилевой пакет
\usepackage{schoolseminar2020}

%\renewcommand{\thefootnote}{*}\footnote{Работа выполнена при частичной финансовой поддержке .....}

\begin{document}
% \udk     - универсальный десятичный классификатор
% \msc     - Индекс предметной классификации (Mathematics Subject Classification)
% \title   - название статьи
% \authors - список авторов
\setcounter{page}{1}

\udk{519.853.4}

\title{
Алгоритм глобальной оптимизации с использованием численных оценок производной минимизируемой функции
\footnote{Исследование выполнено при финансовой поддержке РФФИ в рамках научного проекта № 19-07-00242.}
}

\authors{К.А. Баркалов, А.В. Сысоев, И.С. Ямщиков}
\organizations{Нижегородский государственный университет им. Н.И. Лобачевского}

% \section{название} - заголовок раздела первого уровня
% \subsection{название} - заголовок раздела второго уровня
% \subsubsection{название} - заголовок раздела третьего уровня

\bigskip

В работе рассматривается задача поиска глобального минимума $x^*$ одномерной функции $\varphi(x)$ вида
\begin{equation}\label{problem}
\varphi(x^*)=\min\left\{\varphi(x):x\in\left[a,b\right]\right\}.
\end{equation}
Предполагается, что целевая функция является многоэкстремальной и задана в виде ``черного ящика'', т.е. некоторого алгоритма вычисления ее значений.

В задачах многоэкстремальной оптимизации получение гарантированных оценок глобального минимума возможно только при наличии тех или иных предположений о поведении минимизируемой функции $\varphi(x)$. Одним из наиболее часто используемых предположений является выполнимость условия Липщица
\[
\left|\varphi(x_1)-\varphi(x_2)\right|\leq L\left|x_1-x_2\right|,\; x_1,x_2 \in [a,b],\; 0<L<\infty.
\]
Условие Липщица соответствует предположению ограниченности изменения значения функции при ограниченном изменении ее аргумента. Данное условие позволяет строить оценки возможного поведения функции $\varphi(x)$ на основе конечного множества ее значений, вычисленных в точках области поиска $[a,b]$.

Более сильным предположением о характере поведения функции является липшицевость ее первой производной 
\begin{equation}\label{LipD}
\left|\varphi'(x_1)-\varphi'(x_2)\right|\leq L_1\left|x_1-x_2\right|,\; x_1,x_2 \in [a,b],\; 0<L_1<\infty.
\end{equation}
Выполнение условия (\ref{LipD}) позволяет формировать более точные оценки возможных значений минимизируемой функции $\varphi(x)$, что обеспечивает возможность существенного повышения эффективности разрабатываемых алгоритмов. 
Однако при решении задач оптимизации с функциями вида ``черный ящик'' точные значения производных, как правило, недоступны, а их численная оценка (с помощью операторов численного дифференцирования) требует дополнительных вычислений значений функции. 
При этом даже одно вычисление значения функции (\textit{испытание}) является трудоемкой операцией, т.к. связано с решением задачи численного моделирования.

Следовательно, перспективной является разработка методов глобальной оптимизации, в которых необходимые значения производных вычисляются на основе точек предшествующих итераций, без проведения дополнительных испытаний.

%Предполагается также, что целевая функция является многоэкстремальной и задана в виде ``черного ящика'', т.е. некоторого алгоритма вычисления значения функции и ее производной в зависимости от параметра. При этом каждое \textit{испытание} (т.е. вычисление значений $\varphi(x)$ и $\varphi'(x)$ в точке $x$) является трудоемкой операцией, т.к. связано  с необходимостью численного моделирования.

В работе \cite{Gergel96} был сформулирован алгоритм глобального поиска с использованием производных для решения одномерных задач. 
Данный метод в процессе своей работы предполагает построение последовательности точек $x^k\in[a,b]$, в которых проводятся поисковые испытания. Множество троек $ \{ x^i, \varphi^i=\varphi (x^i), \varphi'^i=\varphi'(x^i)\}, 1\leq i \leq k,$ составляет поисковую информацию, накопленную методом после проведения $k$ шагов.

В рамках проведенного исследования был предложен алгоритм, в котором точные значения производных заменяются их численными оценками, вычисленными на основе ранее накопленной поисковой информации. 
Если ввести нумерацию (нижним индексом) точек проведенных испытаний в порядке возрастания координаты, т.е. $a=x_0<x_1<...<x_k=b$, то в качестве численных оценок значений производных $\overline{\varphi_i}' = \overline{\varphi}'(x_i) \approx \varphi'(x_i)$, $0\leq i\leq k$ можно использовать трехточечную аппроксимацию вида
\[
\overline{\varphi_0}' = \frac{1}{H_1}\left(-(2+\delta_{2})\varphi(x_{0}) + \frac{(\delta_{2}+1)^2}{\delta_{2}}\varphi(x_1) - \frac{1}{\delta_{2}}\varphi(x_{2}) \right),
\]
\begin{equation}\label{numer}
\overline{\varphi_i}' = \frac{1}{H_i}\left(-\delta_{i+1}\varphi(x_{i-1}) - \frac{\delta_{i+1}^2-1}{\delta_{i+1}}\varphi(x_i) + \frac{1}{\delta_{i+1}}\varphi(x_{i+1}) \right),
\end{equation}
\[
\overline{\varphi_k}' = \frac{1}{H_{k-1}}\left(\delta_{k}\varphi(x_{k-2}) - \frac{(\delta_{k}+1)^2}{\delta_{k}}\varphi(x_{k-1}) + \frac{2+\delta_k}{\delta_{k}}\varphi(x_k) \right),
\]
где $H_i=h_i+h_{i+1}$, $\delta_{i+1}=\frac{h_{i+1}}{h_i}$, $h_i=x_i-x_{i-1}$.

Таким образом, новый алгоритм представляет собой дальнейшее развитие метода из \cite{Gergel96}, в котором значения первой производной минимизируемой функции заменяются их численные оценками вида (\ref{numer}).

Для демонстрации эффективности предложенного алгоритма глобального поиска с использованием численных оценок производных (АГПЧП)  было проведено сравнение со следующими методами глобальной оптимизации: метод Гальперина (МГ) \cite{Galperin}, метод Пиявского (МП) \cite{Pijavski72}, алгоритм глобального поиска Стронгина (АГП) \cite{Strongin2013}, алгоритм глобального поиска с использованием производных (АГПП) \cite{Gergel96}. Сравнение было проведено при решении серии из 20 многоэкстремальных задач, использованных для оценки эффективности методов глобальной оптимизации в \cite{Gergel96}.

В проведенных экспериментах точность поиска решения составила $\epsilon=10^{-4}(b-a)$, где $a$ и $b$ -- границы области поиска. Для методов Гальперина и Пиявского использовались точные значения константы Липшица для минимизируемой функции, оцененные предварительно. Для метода Стронгина использовалось значение параметра $r=2$. Для методов с использованием производных использовался параметр $r=1.1$. В таблице 1 отражено среднее число итераций (округленное до целых значений), потребовавшихся методам для решения всех задач серии.

\begin{table}[h]
	\caption{Результаты решения серии задач}
	\begin{center}
		\begin{tabular}{|c|c|c|c|c|c|}
			\hline
			& МГ & МП & МС & АГПП & АГПЧП \\
			\hline
			\hline
			Среднее число итераций & 720 & 314 & 244 & 26 & 24  \\
			\hline
		\end{tabular}
	\end{center}
\end{table}

Результаты экспериментов наглядно демонстрируют превосходство (по числу итераций) методов с использованием производных над методами, в которых производные не используются. Одновременно с этим видно, что методам АГПП и АГПЧП требуется примерно одинаковое число итераций. При этом каждая итерация АГПП подразумевает вычисление точного значения производной, тогда как в АГПЧП никакие дополнительные вычисления значений функции и ее производных не требуются.

Направлением дальнейших исследований будет являться исследование свойств предложенного алгоритма при решении многомерных задач многоэкстремальной оптимизации. Обобщение на многомерные задачи может быть выполнено, например, с использованием схемы диагонального разбиения области поиска \cite{Sergeyev2008} или же с использованием схемы рекурсивной (вложенной) оптимизации \cite{Grishagin2007}.

\begin{biblio}

%\bibitem{Evtushenko2009} Евтушенко Ю.Г., Малкова В.У., Станевичюс А.-И.А. Параллельный поиск глобального экстремума функций многих переменных // Ж. вычисл. матем. и матем. физ. 2009. Т. 49, № 2.  С. 255--269.

%\bibitem{Elsakov} Елсаков С.М., Ширяев В.И. Однородные алгоритмы многоэкстремальной оптимизации // Ж. вычисл. матем. и матем. физ. 2010. Т. 50, № 10. С. 1727--1740.

\bibitem{Gergel96}
Гергель В.П. Об одном способе учета значений производных при минимизации многоэкстремальных функций // Ж. вычисл. матем. и матем. физ. 1996. Т. 36, №5. С. 51--67.

\bibitem{Galperin}
Galperin E.A. The cubic algorithm // J. Math. Anal. Appl. 1985. Vol. 112. P. 635--640.

\bibitem{Pijavski72}
Пиявский С.А. Один алгоритм отыскания абсолютного экстремума функций // Журн. вычисл. матем. и матем. физ. 1972. Т. 12. № 4. С. 888--896.

\bibitem{Strongin2013}
Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А. Параллельные вычисления в задачах глобальной оптимизации. М.: Изд-во МГУ, 2013. 280 с.

\bibitem{Sergeyev2008} Сергеев Я.Д., Квасов Д.Е. Диагональные методы глобальной оптимизации. М.: Физматлит, 2008. 352 с.

\bibitem{Grishagin2007} Городецкий~С.Ю., Гришагин~В.А. Нелинейное программирование и многоэкстремальная оптимизация. Н.~Новгород: Изд-во ННГУ, 2007. 489~с.

%\bibitem{Gergel97} Gergel V.P. A global optimization algorithm for multivariate function with Lipschitzian first derivatives // J. Glob. Optim. 1997. Vol. 10(3). P. 257--281.

\end{biblio}

\msc{90C26}

\title{Global optimization algorithm using numerical estimates of the objective function derivative}

\authors{K.A. Barkalov, A.V. Sysoyev, I.S. Yamshchikov}
\organizations{ Lobachevsky State University of Nizhni Novgorod }

\end{document}

### 2020-07-03 2020/Saransk_Yamshchikov/theses.tex
Add convergence theorem

@@
 
-В задачах многоэкстремальной оптимизации получение гарантированных оценок глобального минимума возможно только при наличии тех или иных предположений о поведении минимизируемой функции $\varphi(x)$.
+В задачах вида (\ref{problem}) получение гарантированных оценок глобального минимума возможно только при наличии тех или иных предположений о поведении минимизируемой функции $\varphi(x)$.
 Одним из наиболее часто используемых предположений является выполнимость условия Липщица
@@
 Однако при решении задач оптимизации с функциями вида ``черный ящик'' точные значения производных, как правило, недоступны, а их численная оценка (с помощью операторов численного дифференцирования) требует дополнительных вычислений значений функции.
-При этом даже одно вычисление значения функции (\textit{испытание}) является трудоемкой операцией, т.к. связано с решением задачи численного моделирования.
+При этом даже одно вычисление значения функции (\textit{испытание}) является трудоемкой операцией, т.к. связано с проведением численного моделирования.
 
-Следовательно, перспективной является разработка методов глобальной оптимизации, в которых необходимые значения производных вычисляются на основе точек предшествующих итераций, без проведения дополнительных испытаний.
+Следовательно, перспективной является разработка методов глобальной оптимизации, в которых необходимые значения производных вычисляются на основе точек предшествующих итераций, без проведения дополнительных поисковых испытаний.
 
@@
 В работе \cite{Gergel96} был сформулирован алгоритм глобального поиска с использованием производных для решения одномерных задач. 
-Данный метод в процессе своей работы предполагает построение последовательности точек $x^k\in[a,b]$, в которых проводятся поисковые испытания.
-Множество троек $ \{ x^i, \varphi^i=\varphi (x^i), \varphi'^i=\varphi'(x^i)\}, 1\leq i \leq k,$ составляет поисковую информацию, накопленную методом после проведения $k$ шагов.
+Данный метод в процессе своей работы предполагает построение последовательности точек $x^k\in[a,b]$, в которых проводятся испытания.
+Множество троек $ \{ x^i, \varphi^i=\varphi (x^i), \varphi'^i=\varphi'(x^i)\}, 0 \leq i \leq k,$ составляет поисковую информацию, накопленную методом после проведения $k$ шагов.
+
+Общая схема алгоритма может быть представлена в следующем виде.
+Первые два испытания проводятся в граничных точках области поиска, т.е.
+$x^0 = a, x^1 = b$.
+Для выбора точки $x^{k+1}, k>1,$ очередного испытания требуется выполнить следующие действия.
+
+\begin{enumerate}
+	\item Перенумеровать точки $x^0,...,x^k$ предшествующих испытаний нижним индексом в порядке возрастания координаты, т.е.
+	\begin{equation}\label{order}
+		a=x_0<x_1<...<x_k=b,
+	\end{equation}
+	и сопоставить им значения $ \varphi_i=\varphi (x_i), \varphi'_i=\varphi'(x_i)\}, 0 \leq i \leq k,$ вычисленные в этих точках.
+	\item Для каждого интервала $(x_{i-1},x_i), 1\leq i\leq k$, вычислить его \textit{характеристику} $R(i)$.
+	\item Определить интервал с максимальной характеристикой 
+	\[
+	R(t)=\max\{R(i), 1\leq i \leq k\}.
+	\]
+	\item Провести следующее испытание в точке интервала с максимальной характеристикой, т.е. $x^{k+1}\in(x_{t-1},x_t).$
+	\item Проверить условие остановки $x_t-x_{t-1}<\epsilon$, где $t$ -- номер интервала с максимальной характеристикой, а $\epsilon > 0$ -- заданная точность. 
+	\end{enumerate}
+
+	Конкретные формулы для вычисления характеристики интервала и точки проведения нового испытания, а также соответствующие теоретические утверждения о сходимости алгоритма представлены в \cite{Gergel96}.
 
 В рамках проведенного исследования был предложен алгоритм, в котором точные значения производных заменяются их численными оценками, вычисленными на основе ранее накопленной поисковой информации. 
-Если ввести нумерацию (нижним индексом) точек проведенных испытаний в порядке возрастания координаты, т.е.
-$a=x_0<x_1<...<x_k=b$, то в качестве численных оценок значений производных $\overline{\varphi_i}' = \overline{\varphi}'(x_i) \approx \varphi'(x_i)$, $0\leq i\leq k$ можно использовать трехточечную аппроксимацию вида
+В соответствии с нумерацией (\ref{order}) точек испытаний в порядке возрастания координаты в качестве численных оценок значений производных $\overline{\varphi_i}' = \overline{\varphi}'(x_i) \approx \varphi'(x_i)$, $0\leq i\leq k$ можно использовать трехточечную аппроксимацию вида
 \[
@@
 \begin{equation}\label{numer}
-\overline{\varphi_i}' = \frac{1}{H_i}\left(-\delta_{i+1}\varphi(x_{i-1}) - \frac{\delta_{i+1}^2-1}{\delta_{i+1}}\varphi(x_i) + \frac{1}{\delta_{i+1}}\varphi(x_{i+1}) \right),
+\overline{\varphi_i}' = \frac{1}{H_i}\left(-\delta_{i+1}\varphi(x_{i-1}) - \frac{\delta_{i+1}^2-1}{\delta_{i+1}}\varphi(x_i) + \frac{1}{\delta_{i+1}}\varphi(x_{i+1}) \right), \; 0<i<k,
 \end{equation}
@@
 Таким образом, новый алгоритм представляет собой дальнейшее развитие метода из \cite{Gergel96}, в котором значения первой производной минимизируемой функции заменяются их численные оценками вида (\ref{numer}).
+Отметим, что в алгоритме вместо априори неизвестного значения константы Липшица $L_1$ для $\varphi'(x)$ используется ее адаптивная оценка $m$, вычисляемая на основе полученной поисковой информации.
+Справедлива следующая теорема о сходимости алгоритма.
+\begin{Teor} Пусть $\{x^k\}$ есть последовательность точек испытаний, порождаемой алгоритмом при минимизации  функции $\varphi(x), x\in[a,b],$ производная которой удовлетворяет условию Липщица с константой $L_1$, и пусть на некотором шаге поиска $k$ для адаптивной оценки константы Липшица $m$ справедливо неравенство
+\[
+rm>\gamma L_1, \; \gamma = \left[ 4(b-a) + (b-a)^2 +0.5 \right]/\epsilon^2 ,
+\]
+где $r>1$ -- параметр, а $\epsilon>0$ -- точность в условии остановки алгоритма.
+Тогда любая точка глобального минимума $x^*$ является предельной точкой последовательности $\{x^k\}$ и, кроме того, любая предельная точка $\overline{x}$ этой последовательности является точкой глобального минимума функции $\varphi(x)$.
+\end{Teor}
 
@@
 В проведенных экспериментах точность поиска решения составила $\epsilon=10^{-4}(b-a)$, где $a$ и $b$ -- границы области поиска.
-Для методов Гальперина и Пиявского использовались точные значения константы Липшица для минимизируемой функции, оцененные предварительно.
-Для метода Стронгина использовалось значение параметра $r=2$.
+Для методов Гальперина и Пиявского использовались точные значения константы Липшица минимизируемой функции, оцененные предварительно.
+Для алгоритма глобального поиска Стронгина использовалось значение параметра $r=2$.
 Для методов с использованием производных использовался параметр $r=1.1$.
@@
 			\hline
-			& МГ & МП & МС & АГПП & АГПЧП \\
+			& МГ & МП & АГП & АГПП & АГПЧП \\
 			\hline

### 2020-07-07 2020/OPTIMA_Sysoyev/samplepaper.tex
OPTIMA: add paper about numerically derivatives

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\usepackage{amsmath}

\begin{document}
%
\title{Global optimization method with numerically calculated function derivatives\thanks{Supported by Russian Foundation for Basic Research (grant 19-07-00242)}}
%
\titlerunning{Global optimization method with numerical derivatives}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Victor Gergel\orcidID{0000-0002-4013-2329} \and Alexander Sysoyev\orcidID{0000-0003-1542-7624}}
	%\inst{1}\orcidID{0000-1111-2222-3333} \and
%Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
%Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{V. Gergel, A. Sysoyev}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhny Novgorod,\\ Nizhny Novgorod, Russian Federation\\
	\email{alexander.sysoyev@itmm.unn.ru, gergel@unn.ru}}
	% \and
%Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
%\email{lncs@springer.com}\\
%\url{http://www.springer.com/gp/computer-science/lncs} \and
%ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
%\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The paper proposes a method for solving computationally time-consuming multidimensional global optimization problems. The developed method combines the use of a nested dimensional reduction scheme, numerical estimates of the objective function derivatives, and optimization schemes for discontinuous functions. Derivatives significantly reduce the cost of solving global optimization problems, however, the use of a nested scheme can lead to the fact that the derivatives of the reduced function become discontinuous. Typical global optimization methods are highly dependent on the continuity of the objective function. Thus, to use derivatives in combination with a nested scheme, an optimization method is required that can work with discontinuous functions. The paper discusses the corresponding method, as well as the results of numerical experiments confirming the applicability of the described approach.

\keywords{Multidimensional optimization \and Global search algorithms \and Lipschitz condition \and Numerical estimations of derivative values \and Dimensionality reduction \and Discontinuous functions \and Numerical experiments.}
\end{abstract}
%
%
%
\section{Introduction}

The global (or multiextremal) unconstrained optimization problem \cite{Floudas1996,Floudas2016,Horst1990,Locatelli2013,Pardalos2016,Paulavicius2014,Pinter1996,Strongin1978,Strongin2000,Strongin2013,Zhigljavsky2008} can be stated as follows
\begin{equation}
\label{eq:problem_statement}
\varphi(y^*) = \min\{\varphi(y)\colon y \in D\},
\end{equation}
where search domain $D$ represents an $N$-dimensional hyperinterval:
\[
D = \{y \in \bbbr^N\colon a_i \leq y_i \leq b_i,\; i = \overline{1,N}\}.
\]

The objective function $\varphi(y)$ is assumed to be a multiextremal one. Also one of the commonly used assumptions is that the minimized function satisfies the Lipschitz condition
\begin{equation}
\label{eq:Lipschitz_for_function}
|\varphi(y_2) - \varphi(y_1)| \leq L \|y_2 - y_1\|,\; y_1, y_2 \in D.
\end{equation} 
where $L > 0$ is the Lipschitz constant, and $\| \cdot \|$ denotes the Euclidean norm in $\bbbr^N$.

The Lipschitz condition corresponds to the assumption of a limited variation of the function value at limited variations of its parameters. This condition allows making the estimates of potential behaviour of the function $\varphi(y)$ based on a finite set of its values computed at some points in the search domain $D$.

To solve problem (1) numerically, optimization methods usually generate a sequence of points $y_k$, which converges to the global optimum $y^*$. The amount of computation can grow exponentially with an increase in the number of variable parameters $N$.

One approach to reduce computational costs is to use differentiability of the objective function. In this case the fulfilment of the Lipschitz condition (2) may be expanded onto the partial derivatives $\varphi_i^\prime(y), 1 \leq i \leq N$ of the objective function as well i.e.
\begin{equation}
\label{eq:Lipschitz_for_derivative}
|\varphi_i^\prime(y_2) - \varphi_i^\prime(y_1)| \leq L_i \|y_2 - y_1\|,\; y_1, y_2 \in D, 1 \leq i \leq N,
\end{equation} 
where $L_i > 0, 1 \leq i \leq N$ are the corresponding Lipschitz constants for the partial derivatives $\varphi_i^\prime(y), 1 \leq i \leq N$~\cite{Baritompa1994,Breiman1993,Gergel1996,Gergel1997,Gergel2017,Lera2013,Sergeyev1998,Sergeyev2015,Shpak1995}.

However, in some applied optimization problems the computing of the derivatives may be restricted or even impossible. In this case the usage of the global optimization methods, in which the necessary values of derivatives are computed numerically may be useful~\cite{Gergel2017,Goryachih2017,Griewank2008,Nocedal2006}.

A widely used approach to solve the problem of multidimensional optimization is to reduce it to one-dimensional. This reduction may be based on using Peano (space-filling) curves~\cite{Piyavskij2006,Strongin2000}, the nested multistep reduction scheme~\cite{Dam2010,Shi2000}, the diagonal generalization technique~\cite{Gergel1997,Pinter1996,Sergeyev2015}, etc. As a result, one-dimensional optimization algorithms can be effectively applied in the multidimensional case~\cite{Baritompa1994,Breiman1993,Dam2010,Galperin1985,Gergel1996,Gergel2017}. It is known that using the nested multistep reduction scheme can lead for executing some redundant global search iterations~\cite{Dam2010,Shi2000,Strongin2000}. This deficiency can be diminished by using the values of derivatives of the objective function -- see the results of numerical experiments given in section 5.

In this paper, the global optimization algorithm utilizing the numerical derivatives of the objective function $\varphi(y)$ is considered. In Section 2, the base one-dimensional algorithm utilizing the numerical derivatives is given. Section 3 introduces a nested dimension reduction scheme that allows one to generalize the proposed one-dimensional algorithm to solving multidimensional global optimization problems. Section 4 describes an approach of usage derivatives in combination with a nested scheme, that can work with discontinuous functions. In Section 5, the results of the numerical experiments are considered, which confirm the developed approach to be promising.

\section{One-dimensional Global Optimization Algorithm Utilizing Numerical	Derivatives}\label{sec:GOAND}

The proposed optimization algorithm is based on the adaptive global method using derivatives (AGMD) ~\cite{Gergel1996,Gergel1997}, designed to solve one-dimensional global optimization problems

\begin{equation}
\label{eq:problem_statement_1d}
\varphi(x^*) = \min\{\varphi(x)\colon x \in [a,b]\}.
\end{equation}

The adaptive global method using numerical derivatives (AGMND) is a modification of AGMD, in which the values of the first derivative of the objective function are replaced by their numerical estimates ~\cite{Gergel1997,Goryachih2017}.

Consider the computational scheme of the AGMND. The first two iterations are performed at the boundary points $a$ and $b$. Then let $k, k > 1$ iterations of the global search were completed, and the values of the objective function $\varphi(x)$ have been computed at each iteration (hereinafter, these computations will be called \textit{trials}). The test point of the next ($k+1$) optimization iteration is determined by the following rules.

\textbf{Rule 1.} Renumber the points of previous trials by subscripts in increasing order

\begin{equation}
a = x_0 < x_1 < \ldots < x_i < \ldots < x_k = b.
\end{equation}

\textbf{Rule 2.} Compute the numerical estimations of the first derivatives of $\varphi(x)$ at the points of the executed search iterations $x_i, 0 \le i \le k$ according to the expressions:
\begin{equation}
\dot{z}_i = \left\lbrace
\begin{array}{lr}
\frac{z_{i+1}-z_i}{x_{i+1}-x_i}, i = 0, \\
\frac{z_i-z_{i-1}}{x_i-x_{i-1}}, 1 \le i \le k,
\end{array}
\right.
\end{equation}
hereinafter $z_i, 0 \le i \le k$ denotes $\varphi(x_i)$.

\textbf{Rule 3.} Compute the estimation of the Lipschitz constant from (\ref{eq:Lipschitz_for_derivative}) for the first derivative of the optimized function
\begin{equation}\label{m}
m = \left\{
\begin{array}{ll}
rM, & M > 0, \\
1, & M = 0,
\end{array}
\right.
\end{equation}
where
\begin{equation}
M = \max(M_{i}), 1 \le i \le k, 
\end{equation}
\begin{equation}
M_{i} = \max \left\{
\begin{array}{ll}
|\dot{z}_i-\dot{z}_{i-1}| / |x_i-x_{i-1}|, \\
-2[z_i-z_{i-1}-\dot{z}_{i-1}(x_i-x_{i-1})] / (x_i-x_{i-1})^2|, \\
2[z_i-z_{i-1}-\dot{z}_i(x_i-x_{i-1})] / (x_i-x_{i-1})^2|,
\end{array}
\right.
\end{equation}
and $r > 1$ is the reliability parameter of the algorithm.

\textbf{Rule 4.} Compute the characteristic $R(i)$ for each interval $(x_{i-1}, x_i), 1 \le i \le k$ according to the following expressions to estimate the minimum possible values of $\varphi(x)$ in the interval ($(x_{i-1}, x_i)$
\begin{equation}
R(i) = \left\{
\begin{array}{ll}
\widehat{\varphi}_i(\widehat{x_i}), & \widehat{x_i} \in [\bar{x}_i,\bar{\bar{x}}_i], \\
\min (\widehat{\varphi}_i(\bar{x}_i),\widehat{\varphi}_i(\bar{\bar{x}}_i)), & \widehat{x_i} \notin [\bar{x}_i,\bar{\bar{x}}_i],
\end{array}
\right.
\end{equation}
where
\begin{equation}
\widehat{x_i} = \frac{-\dot{z}_{i-1}+m(\bar{x}_i-x_{i-1})+mx_i}{m},
\end{equation}
and the auxiliary functions (minorants) $\widehat{\varphi}_i(x), 1 \le i \le k$ take the form
\begin{equation}\label{phi_hat}
\widehat{\varphi}_i(x) = \left\{
\begin{array}{ll}
\widehat{\varphi}_{i1}(x) = z_{i-1}+\dot{z}_{i-1}(x_i-x_{i-1})-0.5m(x-x_{i-1})^2, & x \in (x_{i-1},\bar{x}_i) \\
\widehat{\varphi}_{i2}(x) = A_i(x-\bar{x}_i)+0.5m(x-\bar{x}_i)^2 +B_i, & x \in [\bar{x}_i, \bar{\bar{x}}_i], \\
\widehat{\varphi}_{i3}(x) = z_i-\dot{z}_i(x-x_i)-0.5m(x-x_i)^2, & x \in (\bar{\bar{x}}_i, x_i],
\end{array}
\right.
\end{equation}
where
\begin{equation}
\begin{array}{lr}
A_i=\dot{z}_{i-1}-m(\bar{x}_i-x_{i-1}), \\
B_i = \widehat{\varphi}_{i1}(\bar{x}_i), \\
\bar{x}_i = \frac{(z_{i-1}-\dot{z}_{i-1}x_{i-1})-(z_i-\dot{z}_ix_i)+m(x_i^2-x_{i-1}^2)/2-md_i^2}{m(x_i-x_{i-1})+(\dot{z}_i-\dot{z}_{i-1})} \\
\bar{\bar{x}}_i = \frac{(z_{i-1}-\dot{z}_{i-1}x_{i-1})-(z_i-\dot{z}_ix_i)+m(x_i^2-x_{i-1}^2)/2+md_i^2}{m(x_i-x_{i-1})+(\dot{z}_i-\dot{z}_{i-1})} \\
d_i = (x_i-x_{i-1})/2-(\dot{z}_i-\dot{z}_{i-1})/2m.
\end{array}
\end{equation}

Each characteristic $R(i), 1 \le i \le k$ calculated in this way is an estimation of the minimum possible value of the minorant $\widehat{\varphi}_i(x)$ from (\ref{phi_hat}) in the intervals $[x_{i-1}, x_i]$ and the estimation of the minimum possible values of $\varphi(x)$ in these intervals.

\textbf{Rule 5.} Find the interval $(x_{t-1}, x_t)$ with the minimal characteristic $R(t)$

\begin{equation}\label{R_t}
R(t) = \min \{R(i) : 1 \le i \le k\}.
\end{equation}

In the case when there are several intervals satisfying (\ref{R_t}), for definiteness, the interval with the minimum number $t$ is taken.

\textbf{Rule 6.} Compute the next point of the next trial $x^{k+1}$ accordingly

\begin{equation}\label{key}
x^{k+1} = \left\{
\begin{array}{ll}
\widehat{x}_t, & \widehat{x}_t \in [\bar{x_t},\bar{\bar{x}}_t], \\
\bar{x}_t, & \widehat{\varphi}(\bar{x}_t) \le \widehat{\varphi}(\bar{\bar{x}}_t), \\
\bar{\bar{x}}_t, & \widehat{\varphi}(\bar{x}_t) > \widehat{\varphi}(\bar{\bar{x}}_t).
\end{array}
\right.
\end{equation}

The stopping condition is defined by the following relation
\begin{equation}
|x_t-x_{t-1}| \le \varepsilon,
\end{equation}
where $\varepsilon$ is the accuracy, $\varepsilon > 0$.
The minimum computed value of the objective function is accepted as the current estimate of the global minimum value i.e:

\begin{equation}
\varphi^* = \min \{z_i : 0 \le i \le k\}.
\end{equation}

\textbf{Note 1.} The computing of the numerical estimations $\dot{z}_i, 0 \le i \le k$ of the first derivative of the function $\varphi(x)$ can be performed also using the three-point approximating expressions:
\begin{equation}
\dot{z}_i = \left\{
\begin{array}{ll}
\frac{1}{H_1^2}
\left( -(2+\delta_2)z_0+\frac{(1+\delta_2)^2}{\delta_2}z_1-\frac{1}{\delta_2}z_2 \right), & i = 0, \\
\frac{1}{H_i^{i+1}}
\left( -\delta_{i+1}z_{i-1}+\frac{\delta_{i+1}^2-1}{\delta_{i+1}}z_i+\frac{1}{\delta_{i+1}}z_{i+1}
\right), & 1 < i < k, \\
\frac{1}{H_{k-1}^k}
\left(
\delta_k z_{k-2}-\frac{(1+\delta_k)^2}{\delta_k}z_{k-1}+\frac{(2\delta_k+1)}{\delta_k}z_k
\right), & i = k,
\end{array}
\right.
\end{equation}
where $H_i^{i+1} = h_i+h_{i+1}, \delta_{i+1} =\frac{h_{i+1}}{h_i}$ and $h_i=x_i-x_{i-1}$, see~\cite{Griewank2008}.
These formula used three points of trials, and can be used in the \textbf{Rule 2} if $k > 2$.

\textbf{Note 2.} For the applicability of the computational scheme described above, the fulfilment of the following inequalities
\begin{equation}\label{note2}
x_{i-1} < \bar{x}_i < \bar{\bar{x}}_i < x_i
\end{equation}
for all intervals $(x_{i-1}, x_i), 1 \le i \le k$ is necessary. If the estimate of the Lipschitz constant $m$ computed in (\ref{m}) is insufficient and the condition (\ref{note2}) is violated for some $1 \le i \le k$ the value $m$ should be refined. Thus, the maximum root of the equations
\begin{equation}
\left\{
\begin{array}{ll}
-(z_i-z_{i-1})+0.5(\dot{z}_i+\dot{z}_{i-1})+0.25m(x_i-x_{i-1})^2-\frac{(\dot{z}_i-\dot{z}_{i-1})^2}{4m}=0, \\
(z_i-z_{i-1})-0.5(\dot{z}_i+\dot{z}_{i-1})+0.25m(x_i-x_{i-1})^2-\frac{(\dot{z}_i-\dot{z}_{i-1})^2}{4m}=0 ,
\end{array}
\right.
\end{equation}
was selected as $m$ in this case for AGMD~\cite{Gergel1996}.

\section{Nested Dimensionality Reduction Scheme}\label{sec:nestedreduction}

One approach to solve multidimensional optimization problems is dimension reduction, which allows to use effective one-dimensional optimization methods. In this paper, dimension reduction is performed using the well-known nested scheme~\cite{Dam2010,Shi2000,Strongin1978,Strongin2000,Strongin2013}. According to this scheme, the solving of a multidimensional optimization problem (\ref{eq:problem_statement}) can be obtained by solving a series of nested one-dimensional problems:

\begin{equation}
\min \{ \varphi(y) : y \in D\} = \min_{[a_1,b_1]} \ldots \min_{[a_N,b_N]} \varphi(y_1, \ldots, y_N).
\end{equation}

In other words the solving of problem (\ref{eq:problem_statement}) is reduced to solving a one-dimensional problem:
\begin{equation}\label{eq:nested}
\varphi(y^*) = \min_{y \in D} = \min_{y_1 \in [a_1,b_1]}\widetilde{\varphi}_1(y_1),
\end{equation}
where
\begin{equation}\label{eq:nested1}
\begin{array}{c}
\widetilde{\varphi}_i(y_i)=\varphi_i(y_1, \ldots, y_i) = \min\limits_{y_{i+1} \in [a_{i+1},b_{i+1}]}\varphi_i(y_1, \ldots, y_i,y_{i+1}), 1 \le i \le N, \\ \\
\widetilde{\varphi}_N(y_1, \ldots, y_N) = \varphi(y_1, \ldots, y_N).
\end{array}
\end{equation}

The one-dimensional function in (\ref{eq:nested}) is constructed according to a general recursive scheme -- in order to compute the values $\widetilde{\varphi}_1(y_1)$ for some given value of the variable
$y_1 = \widehat{y_1}$ it is necessary to minimize the function

\begin{equation}
\widetilde{\varphi}_2(y_2) = \varphi_2(\widehat{y}_1,y_2).
\end{equation}

With respect to $y_2$ the function $\widetilde{\varphi}_2(y_2)$ is a one-dimensional one as well since the value of the variable $y_1$ is given and fixed one. Next, in turn, in order to compute the value of $\widetilde{\varphi}_2(y_2)$ at the point $y_2 = \widehat{y2}$, it is necessary to minimize the function
\begin{equation}
\widetilde{\varphi}_3(y_3) = \varphi_3(\widehat{y}_1,\widehat{y}_2,y_3),
\end{equation}
and so forth.

Additional information on the nested dimensionality reduction scheme and its applications for solving the multidimensional global optimization problems can be found, for example, in~\cite{Dam2010,Shi2000,Strongin2000,Strongin2013}.

\section{Global Optimization Algorithm for Discontinuous Functions}\label{sec:GOADF}

To solve problem (\ref{eq:problem_statement}), the method with derivatives can only be used if the objective function is smooth. But one-dimensional functions $\widetilde{\varphi}_i(y_i), 1 \le i < N$ from (\ref{eq:nested1}) (except the function of the
last decomposition level $\widetilde{\varphi}_N(y_N)$) in the nested reduction scheme can be non-smooth at some points i.e. the derivatives of these functions can be discontinuous at these points.

Strictly speaking, the AGMND method described in Section 2 can only be used to minimize one-dimensional function $\widetilde{\varphi}_N(y_N)$) at the last level of decomposition. The paper~\cite{Gergel2019} presents the results of experiments comparing a number of optimization methods, which showed that the AGMND method provides good efficiency even in the case of nonsmooth reduced functions. But with increasing the problem dimensionality ($N$) the number of discontinuity points of the derivatives can grow exponentially and increase the number of executed trials.

To handle the discontinuity problem one can use the combined method which uses AGMND to minimize the one-dimensional function $\widetilde{\varphi}_N(y_N)$) and any one-dimensional method without derivatives for the remaining functions $\widetilde{\varphi}_i(y_i), 1 \le i < N$.

Another way is to use an optimization method that can work with discontinuous functions. As such a method, the paper considers the modified Strongin algorithm~\cite{Strongin1990}. Consider its computational scheme.

The first iteration is performed at any point in the interval $(a, b)$. Then let $k, k > 1$ iterations of the global search were completed. The test point of the next ($k+1$) optimization iteration is determined by the following rules.

\textbf{Rule 1.} Renumber the points of previous trials by subscripts in increasing order

\begin{equation}
a = x_0 < x_1 < \ldots < x_i < \ldots < x_k = b.
\end{equation}

\textbf{Rule 2.} Compute

\begin{equation}
\mu_i = \frac{|z_i-z_{i-1}|}{|x_i-x_{i-1}|}, 1 \le i \le k.
\end{equation}

\textbf{Rule 3.} Reorder $\mu_i$ values in descending order

\begin{equation}
\mu(1) > \mu(2) > \ldots > \mu(k)
\end{equation}
and determine a minimal number $p$ such that

\begin{equation}\label{eq:p}
\frac{\mu(p)}{\mu(p+1)} \ge Q, 1 \le p < qk,
\end{equation}
where $q$ and $Q$ are given numbers ($0 < q < 1 < Q$).

Construct a subset of numbers

\begin{equation}
J = \{i: 1 \le i \le k, \mu_i = \mu(j), 1 \le j \le p \}
\end{equation}
of those intervals to which sufficiently large $\mu_i$ correspond.

\textbf{Rule 4.} Set for all intervals

\begin{equation}
\delta_i = \left\{
\begin{array}{ll}
sign(z_i-z_{i-1}), & i \in J, \\
0, & i \notin J,
\end{array}
\right. 1 \le i \le k.
\end{equation}

\textbf{Rule 5.} Compute the estimation of the Lipschitz constant

\begin{equation}
\mu = \max(\mu_{i}), 1 \le i \le k
\end{equation}
and the characteristic $R(i)$ for each interval $(x_{i-1}, x_i), 1 \le i \le k$

\begin{equation}
R(i) = (1+|\delta_i|)\Delta_i+(1-|\delta_i|)\frac{(z_i-z_{i-1})^2}{(r\mu)^2\Delta_i}-2\frac{(1-\delta_i)z_i+(1+\delta_i)z_{i-1}}{r\mu},
\end{equation}
where $\Delta_i = x_i - x_{i-1}$.

\textbf{Rule 6.} Find the interval $(x_{t-1}, x_t)$ with the maximal characteristic $R(t)$

\begin{equation}
R(t) = \max \{R(i) : 1 \le i \le k\}.
\end{equation}

\textbf{Rule 7.} Compute the next point of the next trial $x^{k+1}$ accordingly

\begin{equation}
x^{k+1} = \frac{x_t+x_{t-1}}{2}-(1-|\delta_t|)\frac{z_t-z_{t-1}}{2r\mu}.
\end{equation}

\section{Results of Numerical Experiments}\label{sec:results}

Paper~\cite{Gergel2019} presents the results of comparing the Brent Algorithm (BA)~\cite{Brent1973}, the Strongin Algorithm (GSA)~\cite{Strongin1978,Strongin2000}, GSA algorithm with the dimensionality reduction using Peano space-filling curves~\cite{Strongin1978} (GSA-Peano), AGMD, AGMND and two combined methods (GSA-D and GSA-ND). In the last two methods, GSA was applied to optimize the reduced one-dimensional functions $\widetilde{\varphi}_1(y_1)$ from (\ref{eq:nested1}).

A family of well-known two-dimensional multiextremal test functions~\cite{Strongin1978,Strongin2000,Strongin2013} defined by the relations:
\begin{equation}\label{eq:grish}
\begin{gathered}
\varphi(y_1,y_2)=-\left\{
\Bigl(\sum_{i=1}^7\sum_{j=1}^7\left[A_{ij}a_{ij}(y_1,y_2)+B_{ij}b_{ij}(y_1,y_2)\right]\Bigr)^2+\right.\\
\left.+\Bigl(\sum_{i=1}^7\sum_{j=1}^7\left[C_{ij}a_{ij}(y_1,y_2)+D_{ij}b_{ij}(y_1,y_2)\right]\Bigr)^2\right\}^{\frac{1}{2}},\\
a_{ij}(y_1,y_2)=\sin(\pi i y_1)\sin(\pi j y_2),\\
b_{ij}(y_1,y_2)=\cos(\pi i y_1)\cos(\pi j y_2),
\end{gathered}
\end{equation}
where $0 \le y1, y2 \le 1$, were used. The values $-1 \le A_{ij}, B_{ij}, C_{ij}, D_{ij} \le 1$ are the independent random generated parameters distributed uniformly over the interval $[-1, 1]$.

In all these methods, an adaptive scheme of computing the reliability parameter $r$ from (\ref{m}) was applied: $r=3+10/k$ where $k$ is the number of executed trials. The accuracy of solving the optimization problems was $\varepsilon = 0.001$. In order to obtain the reliable
results, 100 multiextremal problems (\ref{eq:grish}) were solved in every series of experiments.

These results demonstrated the AGMND method to show the best performance (the smallest number of trials).

Thus, to evaluate the efficiency of the scheme described in the section \ref{sec:GOADF}, further experiments were carried out only with the AGMND and GSA-ND methods. In the last method, the modified Strongin algorithm was applied to optimize the reduced one-dimensional functions $\widetilde{\varphi}_1(y_1)$ from (\ref{eq:nested1}).

From the experimental results, the operational characteristics of the compared methods were constructed. The operational characteristic is a graph of the number of solved problems (the ordinate axis) vs the number of executed trials (the abscissa axis)~\cite{Strongin1978,Strongin2000,Strongin2013}. The operational characteristics of the compared methods are presented in Fig.~\ref{fig1}.

\begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{fig1.png}
	\caption{Operational characteristics of the compared optimization methods. The vertical axis is the percentage of problems solved with the required accuracy, the horizontal axis is the number of executed trials} \label{fig1}
\end{figure}

As shown GSA-ND with discontinuous modification achieved $100\%$ solvability later than AGMND, but GSA-ND solved $75\%$ of problems faster than AGMND. The results of the numerical experiments allow us to conclude that the proposed approach is quite promising, but additional studies are necessary.

\section*{Conclusion}\label{sec:conclusion}

In this paper, an efficient method for solving computationally time-consuming multidimensional global optimization problems is proposed.

The developed method combines nested dimension reduction scheme, numerical estimation of the derivatives of the objective function, and discontinuous function optimization schemes.

The developed method combines the use of a nested dimensional reduction scheme, numerical estimates of the objective function derivatives, and optimization schemes for discontinuous functions. Derivatives significantly reduce the cost of solving global optimization problems, however, the use of a nested scheme can lead to the fact that the derivatives of the reduced function become discontinuous. Typical global optimization methods are highly dependent on the continuity of the objective function. Thus, to use derivatives in combination with a nested scheme, an optimization method is required that can work with discontinuous functions. The paper discusses the corresponding method, as well as the results of numerical experiments confirming the applicability of the described approach.

%
\begin{thebibliography}{8}

\bibitem{Baritompa1994}
Baritompa W.: Accelerations for a variety of global optimization methods. Journal of Global Optimization, \textbf{4} 37--45 (1994).

\bibitem{Breiman1993}
Breiman L. and Cutler A.: A deterministic algorithm for global optimization. Mathematical Programming, \textbf{58} 179--199 (1993).

\bibitem{Brent1973}
Brent R.P.: Algorithms for Minimization Without Derivatives. Englewood Cliffs NJ: Prentice-Hall (1973).

\bibitem{Dam2010}
Dam E.R., Husslage B. and Hertog D.: One-dimensional nested maximin designs, Journal of Global Optimization \textbf{46}, 287--306 (2010).

\bibitem{Floudas1996}
Floudas C.A. and Pardalos M.P.: State of the Art in Global Optimization. Computational Methods and Applications. Dordrecht: Kluwer Academic Publishers (1996).

\bibitem{Floudas2016}
Floudas C.A. and Pardalos M.P.: Recent Advances in Global Optimization. Princeton University Press (2016).

\bibitem{Galperin1985}
Galperin E.A.: The cubic algorithm. Journal of Mathematical Analysis and Applications \textbf{112}, 635--640 (1985).

\bibitem{Gergel1996}
Gergel V.P.: A method of using derivatives in the minimization of multiextremum functions. Computational Mathematics and Mathematical Physics, \textbf{36} 729--742 (1996) (In Russian).

\bibitem{Gergel1997}
Gergel V.P.: A global optimization algorithm for multivariate function with Lipschitzian first derivatives. Journal of Global Optimization, \textbf{10}  257--281 (1997).

\bibitem{Gergel2017}
Gergel V.P. and Goryachih A.S.: Global optimization using numerical approximations of derivatives, Learning and Intelligent Optimization. In: LION 2017. Lecture Notes in Computer Science 10556, 320--325 (2017).

\bibitem{Gergel2019}
Gergel V. and Goryachih A.: Multidimensional global optimization using numerical estimates of objective function derivatives. In: Optimization Methods and Software (2019).

\bibitem{Goryachih2017}
Goryachih A.S. and Rachinskaya M.A.: Multidimensional global optimization method using numerically calculated derivatives. Procedia Computer Science \textbf{119}, 90--96 (2017).

\bibitem{Griewank2008}
Griewank A. and Walther A.: Evaluating Derivatives: Principles and Techniques of Algorithmic Differentiation. SIAM (2008).

\bibitem{Horst1990}
Horst R. and Tuy H.: Global Optimization: Deterministic Approaches. Springer-Verlag (1990).

\bibitem{Lera2013}
Lera D. and Sergeyev Y.D.: Acceleration of univariate global optimization algorithms working with Lipschitz functions and Lipschitz first derivatives. SIAM Journal on Optimization, \textbf{23}, 508--529 (2013).

\bibitem{Locatelli2013}
Locatelli M. and Schoen F.: Global Optimization: Theory, Algorithms, and Applications. SIAM (2013).

\bibitem{Nocedal2006}
Nocedal J. and Wright S.: Numerical Optimization. Springer (2006).

\bibitem{Pardalos2016}
Pardalos M.P., Zhigljavsky A.A. and \u{Z}ilinskas J., Advances in Stochastic and Deterministic
Global Optimization, Springer, 2016.

\bibitem{Paulavicius2014}
Paulavi\u{c}ius R. and \u{Z}ilinskas J.: Simplicial Global Optimization. Springer Briefs in Optimization. Springer (2014).

\bibitem{Pinter1996}
Pint\'{e}r J.D.: Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications). Kluwer Academic Publishers (1996).

\bibitem{Piyavskij2006}
Piyavskij S.: An algorithm for finding the absolute extremum of a function. Computational Mathematics and Mathematical Physics \textbf{12}, 57--67 (1972) (In Russian).

\bibitem{Sergeyev1998}
Sergeyev Y.D.: Global one-dimensional optimization using smooth auxiliary functions. Mathematical Programming \textbf{81}, 127--146 (1998).

\bibitem{Sergeyev2015}
Sergeyev Y.D.: A deterministic global optimization using smooth diagonal auxiliary functions. Communications in Nonlinear Science and Numerical Simulation \textbf{21}, 99--111 (2015).

\bibitem{Shi2000}
Shi L. and \'{O}lafsson S.: Nested partitions method for global optimization. Operations Research \textbf{48}, 390--407 (2000).

\bibitem{Shpak1995}
Shpak A.: Global optimization in one-dimensional case using analytically defined derivatives of objective function. Computer Science Journal of Moldova \textbf{3}, 168--184 (1995).

\bibitem{Strongin1978}
Strongin R.G.: Numerical Methods in the Multiextremal Problems (information-statistical algorithms). Nauka (1978) (In Russian).

\bibitem{Strongin1990}
Strongin R.G.: Search of global optimum. Znanie (1990) (In Russian).

\bibitem{Strongin2000}
Strongin, R.G., Sergeyev, Ya.D.: Global Optimization with non-convex constraints: sequential and parallel algorithms. Dordrecht: Kluwer Academic Publishers. (2000, 2nd ed. 2013, 3rd ed. 2014)

\bibitem{Strongin2013}
Strongin R.G., Gergel V.P., Grishagin V.A. and Barkalov K.A.: Parallel Computations in the Global Optimization Problems. MSU Publishing (2013) (In Russian).

\bibitem{Zhigljavsky2008}
Zhigljavsky A. and \u{Z}ilinskas A.: Stochastic Global Optimization. Springer, Berlin (2008).

\end{thebibliography}

\end{document}

### 2020-07-13 2020/Saransk_Usova/theses.tex
Theses to Saransk - second version with final experiments results

@@
 
-Задачи одномерной оптимизации играют важную роль, т.к.
-многие подходы к решению многомерных задач оптимизации так или иначе основаны на сведении решения исходной многомерной задачи к решению серии связанных задач одномерной оптимизации (см., например, \cite{Grishagin2007,Sergeyev2008}).
+Стоит отметить, что задачи одномерной оптимизации, несмотря на свою простоту, играют важную роль, поскольку многие подходы к решению многомерных задач оптимизации так или иначе основаны на сведении решения исходной многомерной задачи к решению серии связанных задач одномерной оптимизации (см., например, \cite{Grishagin2007,Sergeyev2008}).
 В литературе описаны различные методы решения задачи (\ref{problem}) в зависимости от предположений о свойствах целевой функции.
-Одним из подобных предположений, которое часто выполняется в прикладных оптимизационных задачах, является предположение о том, что целевая функция $\varphi(x)$ удовлетворяет условию Липшица с априори неизвестной константой $L$
+Одним из подобных наложенных условий на вид целевой функции, которое часто выполняется в прикладных оптимизационных задачах, является предположение о том, что целевая функция $\varphi(x)$ удовлетворяет условию Липшица с априори неизвестной константой $L$
 \[
@@
 Однако в некоторых прикладных задачах свойство липшицевости может не выполняться в силу наличия скачкообразных изменений значений функции в ряде точек области поиска.
+Такие изменения в поведении целевой функции могут являться отображением в математической модели специфики оптимизируемого объекта (например, скачков геометрических размеров или свойств материала, ударных воздействий, резонансных явлений и т.п.).
 Мы будем интерпретировать подобные резкие изменения как разрывы первого рода.
 Иногда множество точек, в которых характеристики объекта претерпевают скачок, известно заранее.
-Вместе с тем существуют задачи, в которых нет априорных оценок точек разрыва, но есть информация, что такие точки возможны.
+Вместе с тем существуют задачи, в которых нет априорных оценок точек разрыва, но имеется информация, что такие точки возможны.
 
@@
 
-В новом алгоритме глобального поиска для разрывных функций (АГП-Р) учет разрывов целевой функции происходит явно, т.е. без операции сглаживания разрыва. 
+В новом алгоритме глобального поиска для разрывных функций (АГП-Р) учет разрывов целевой функции происходит явно, т.е.
+без решения вспомогательной задачи пострения  сглаживающего преобразования и самой операции сглаживания разрыва.
 При этом охватываются случаи как заданных, так и не заданных точек разрыва: при отсутствии информации о координатах точек разрыва их оценка проводится адаптивно, на основе вычисленных значений функции в точках проведенных испытаний.
-Ошибочное определение некоторых точек разрыва не приводит к потере сходимости метода.
+Вместо трудоёмкой операции сглаживания в данном алгоритме используется адаптированная версия вычислительных формул \textit{алгоритма глобального поиска}, которая при построении оценок неизвестной константы $L$, а также вычислении характеристик интервалов и новых точек испытаний исключает вклад точек, идентифицированных как подозрительные на разрыв, или априори известных точек разрыва.
+
+Ошибочное или частично верное определение некоторых точек разрыва не приводит к потере сходимости метода.
 Решающие правила алгоритма интерпретируют такие точки как случай устранимого разрыва, поскольку левый и правый пределы оптимизируемой функции в таких точках совпадают.
@@
 В качестве примера рассмотрим задачу минимизации функции
+
 \begin{equation}\label{test}
@@
 
-Для решения задачи использовался алгоритм глобального поиска для разрывных функций (АГП-Р), точность поиска решения была выбрана $\epsilon = 0.001$.
+Для решения задачи использовался алгоритм глобального поиска для разрывных функций (АГП-Р), точность поиска решения была выбрана равной $\epsilon = 0.001$.
+Параметры самого алгоритма были выставлены по умолчанию.
 На рис.
 \ref{ris1} изображен график функции, на котором наглядно видны точки разрывов.
-Штрихами под графиком обозначены точки поисковых испытаний, потребовавшихся алгоритму для решения задачи с указанной точностью. 
-В случае, когда точки разрывов были известны, метод провел XXX испытаний (штрихи верхнего ряда).
-В случае, когда точки разрывов не были заданы заранее и идентифицировались в процессе поиска, методу потребовалось YYY испытаний (штрихи нижнего ряда).
-В обоих случаях распределение точек испытаний показывает наличие сходимости лишь к глобальному экстремуму, накопления точек около локальных экстремумов --- отсутствуют.
+Штрихами под графиком обозначены точки поисковых испытаний, потребовавшихся алгоритму для решения задачи с заданной точностью.
 
-\begin{figure}%[!ht]
+\begin{figure}[!ht]
 	\begin{center}
-			\includegraphics[width=0.7\linewidth]{ris1.png}
+			\includegraphics[width=0.7\linewidth]{ris1_2.png}
 			\caption{График разрывной целевой функции и точки испытаний} %% подпись к рисунку
@@
 
+В случае, когда точки разрыва были известны, метод провел 74 испытания (штрихи верхнего ряда).
+В случае, когда все точки разрыва не были заданы заранее и идентифицировались в процессе поиска, методу потребовалось 140 испытаний (штрихи нижнего ряда).
+В обоих случаях распределение точек испытаний показывает наличие сходимости лишь к глобальному экстремуму, накопления точек около локальных экстремумов --- отсутствуют.
+Стоит также отметить, что несмотря на то, что в случае незаданных точек разрыва число итераций метода увеличилось в 2 раза, алгоритм избегал проведения итераций в окрестностях точек разрыва благодаря правильной идентификации этих точек.
+
 Для демонстрации эффективности алгоритма глобального поиска для разрывных функций (АГП-Р), проведем его сравнение с алгоритмом имитации отжига (ИО), реализованным в MATLAB Global Optimization Toolbox.
-В качестве основного критерия сравнения будем использовать число поисковых испытаний (т.е.
-вычислений значений целевой функции), выполненное методом до достижения сходимости.
+Итерация данного алгоритма начинается с генерирования новой случайной точки с использованием датчика псевдослучайных чисел.
+Расстояние между новой и текущей точкой определяется по распределению вероятностей, шкала которого пропорциональна текущей температуре (график изменения температуры, т.е.
+отжига, является параметром алгоритма).
+Если целевая функция в новой точке лучше, чем в текущей, то новая точка становится текущей, в противном случае новая точка может стать текущей только с некоторой вероятностью.
+Этот подход и позволяет избегать ``застревания'' в локальном минимуме, что делает метод применимым к решению задач многоэкстремальной оптимизации.
+Помимо этого ИО заявляется как метод, позволяющий решать задачи в том числе и с разрывами, что позволяет использовать его для сравнения с реализованным АГП-Р.
 
-Сравнение алгоритмов проводилось при решении 100 задач с разрывной целевой функцией, которые были построены аналогично задаче (\ref{test}).
-В таблице 1 представлено число решенных задач и среднее число испытаний, выполненных методами при решении серии.
-Задача считалась решенной, если для точки некоторого испытания $x^k$ выполнялось условие $\left|x^k-x^*\right|\leq \delta$, где $x^*$ --- известное решение задачи, а $\delta = 10^{-2}$.
-При этом точность, используемая в условии остановки методов, была выбрана на порядок меньшей, $\epsilon=10^{-3}$.
+В качестве основного критерия сравнения будем использовать число поисковых испытаний (т.е. вычислений значений целевой функции), выполненное методом до достижения сходимости.
+
+Сравнение алгоритмов проводилось при решении серии из 100 задач с разрывной целевой функцией, которые были построены аналогично задаче (\ref{test}), а именно сгенерированные случайным образом непрервыные периодические функции вида (\ref{seria})
+\begin{equation}\label{seria}
+\varphi(x) = \sum_{i=0}^4{A_i\sin(5\pi i x)) + B_i\cos(5\pi i x))},
+\end{equation}
+где $A_i$, $B_i$ – случайно заданные коэффициенты, со значениями из интервала $[-6,6]$, а $x\in[0,1]$, были дополнены четырьмя точками разрыва со случайным образом определенным положением точки, а также значением скачка из диапазона $[-20,30]$.
+
+В Таблице 1 представлено число решенных задач и среднее число испытаний, выполненных сравниваемыми методами при решении серии.
+Задача считалась решенной, если для точки некоторого испытания $x^k$ выполнялось условие $\left|x^k-x^*\right|\leq \epsilon$, где $x^*$ --- известное решение задачи, а $\epsilon=10^{-3}$ --- точность, используемая в условии остановки методов.
 
@@
 			\hline
-			ИО    &  89 &   876  \\
+			ИО    &  92 &   758  \\
 			\hline
-			АГП-Р & 100  & 81 \\
+			АГП-Р & 100  & 76 \\
 			\hline
@@
 
-Результаты показывают, что метод АГП-Р успешно решил все задачи серии, тогда как метод имитации отжига не справился с решением 11 задач, затратив при этом в среднем в 10 раз больше испытаний.
+Результаты показывают, что метод АГП-Р успешно решил все задачи серии, тогда как метод имитации отжига не справился с решением 8 задач, затратив при этом в среднем в 10 раз больше испытаний.
 

### 2020-07-14 2020/Saransk_Usova/theses.tex
some minor corrections

@@
 Решение задач глобальной оптимизации с учетом возможных разрывов целевой функции
-\footnote{Исследование выполнено при финансовой поддержке РФФИ в рамках научного проекта № 19-07-00242.}
+\footnote{Исследование выполнено при финансовой поддержке РФФИ (проект № 19-07-00242) и научно-образовательного математического центра (075-02-2020-1483/1)}
 }
@@
 В докладе на конференции будет представлен разработанный в ННГУ им. Н.И. Лобачевского эффективный алгоритм для поиска \textit{глобального решения} многоэкстремальных задач с разрывными функциями. 
-Данный алгоритм является дальнейшим развитием \textit{алгоритма глобального поиска}, описание, теория сходимости и различные модификации которого представлены в \cite{Strongin2013}.
+Данный алгоритм является дальнейшим развитием \textit{алгоритма глобального поиска} (АГП), описание, теория сходимости и различные модификации которого представлены в \cite{Strongin2013}.
 
 В новом алгоритме глобального поиска для разрывных функций (АГП-Р) учет разрывов целевой функции происходит явно, т.е.
-без решения вспомогательной задачи пострения  сглаживающего преобразования и самой операции сглаживания разрыва.
+без решения вспомогательной задачи построения сглаживающего преобразования и самой операции сглаживания разрыва.
 При этом охватываются случаи как заданных, так и не заданных точек разрыва: при отсутствии информации о координатах точек разрыва их оценка проводится адаптивно, на основе вычисленных значений функции в точках проведенных испытаний.
-Вместо трудоёмкой операции сглаживания в данном алгоритме используется адаптированная версия вычислительных формул \textit{алгоритма глобального поиска}, которая при построении оценок неизвестной константы $L$, а также вычислении характеристик интервалов и новых точек испытаний исключает вклад точек, идентифицированных как подозрительные на разрыв, или априори известных точек разрыва.
 
-Ошибочное или частично верное определение некоторых точек разрыва не приводит к потере сходимости метода.
+Вместо трудоёмкой операции сглаживания в АГП-Р при построении оценок неизвестной константы Липшица $L$ целевой функции проводится приближенная идентификация точек, подозрительных на разрыв.
+Ошибочное определение некоторых точек разрыва не приводит к потере сходимости метода.
 Решающие правила алгоритма интерпретируют такие точки как случай устранимого разрыва, поскольку левый и правый пределы оптимизируемой функции в таких точках совпадают.
@@
 Для решения задачи использовался алгоритм глобального поиска для разрывных функций (АГП-Р), точность поиска решения была выбрана равной $\epsilon = 0.001$.
-Параметры самого алгоритма были выставлены по умолчанию.
+%Параметры самого алгоритма были выставлены по умолчанию.
 На рис.
@@
 
-Для демонстрации эффективности алгоритма глобального поиска для разрывных функций (АГП-Р), проведем его сравнение с алгоритмом имитации отжига (ИО), реализованным в MATLAB Global Optimization Toolbox.
-Итерация данного алгоритма начинается с генерирования новой случайной точки с использованием датчика псевдослучайных чисел.
+Для демонстрации эффективности АГП-Р проведем его сравнение с алгоритмом имитации отжига (ИО), реализованным в MATLAB Global Optimization Toolbox. 
+%Итерация данного алгоритма начинается с генерирования новой точки с использованием датчика псевдослучайных чисел.
 Расстояние между новой и текущей точкой определяется по распределению вероятностей, шкала которого пропорциональна текущей температуре (график изменения температуры, т.е.
@@
 Если целевая функция в новой точке лучше, чем в текущей, то новая точка становится текущей, в противном случае новая точка может стать текущей только с некоторой вероятностью.
-Этот подход и позволяет избегать ``застревания'' в локальном минимуме, что делает метод применимым к решению задач многоэкстремальной оптимизации.
+Этот подход и позволяет избегать <<застревания>> в локальном минимуме, что делает метод применимым к решению задач многоэкстремальной оптимизации.
+Помимо этого ИО заявляется как метод, позволяющий решать задачи в том числе и с разрывами, что позволяет использовать его для сравнения с реализованным АГП-Р.
+Итерация данного алгоритма начинается с генерирования новой точки с использованием датчика псевдослучайных чисел.
+Расстояние между новой и текущей точкой определяется пропорционально текущей <<температуре>> (т.е.
+значению функции в точке).
+Если целевая функция в новой точке лучше, чем в текущей, то новая точка становится текущей; в противном случае новая точка может стать текущей только с некоторой вероятностью.
+Этот подход и позволяет избегать <<застревания>> в локальном минимуме, что делает метод применимым к решению задач многоэкстремальной оптимизации.
 Помимо этого ИО заявляется как метод, позволяющий решать задачи в том числе и с разрывами, что позволяет использовать его для сравнения с реализованным АГП-Р.
@@
 В качестве основного критерия сравнения будем использовать число поисковых испытаний (т.е. вычислений значений целевой функции), выполненное методом до достижения сходимости. 
-
-Сравнение алгоритмов проводилось при решении серии из 100 задач с разрывной целевой функцией, которые были построены аналогично задаче (\ref{test}), а именно сгенерированные случайным образом непрервыные периодические функции вида (\ref{seria})
-\begin{equation}\label{seria}
-\varphi(x) = \sum_{i=0}^4{A_i\sin(5\pi i x)) + B_i\cos(5\pi i x))},
-\end{equation}
-где $A_i$, $B_i$ – случайно заданные коэффициенты, со значениями из интервала $[-6,6]$, а $x\in[0,1]$, были дополнены четырьмя точками разрыва со случайным образом определенным положением точки, а также значением скачка из диапазона $[-20,30]$.
+Сравнение алгоритмов проводилось при решении серии из 100 задач с разрывной целевой функцией, которые были построены аналогично задаче (\ref{test}).
+Сгенерированные случайным образом непрерывные периодические функции
+\[
+\varphi(x) = \sum_{i=0}^4{ \left[A_i\sin(5\pi i x) + B_i\cos(5\pi i x)\right] }, \; x\in[0,1],
+\]
+где коэффициенты $A_i$, $B_i$ выбирались случайно и равномерно из интервала $[-6,6]$, были дополнены четырьмя точками разрыва (выбранными случайно из [0,1]) и значением скачка из диапазона $[-20,30]$.
 

### 2020-08-11 2020/Saransk_Usova/theses.tex
fix grant numbers

@@
 Решение задач глобальной оптимизации с учетом возможных разрывов целевой функции
-\footnote{Исследование выполнено при финансовой поддержке РФФИ (проект № 19-07-00242) и научно-образовательного математического центра (075-02-2020-1483/1)}
+\footnote{Исследование выполнено при финансовой поддержке РФФИ (проект 19-07-00242), научно-образовательного математического центра (075-02-2020-1483/1) и государственного задания (0729-2020-0055).}
 }

### 2020-08-11 2020/Saransk_Yamshchikov/theses.tex
fix grant numbers

@@
 Алгоритм глобальной оптимизации с использованием численных оценок производной минимизируемой функции
-\footnote{Исследование выполнено при финансовой поддержке РФФИ в рамках научного проекта № 19-07-00242.}
+\footnote{Исследование выполнено при финансовой поддержке РФФИ (проект 19-07-00242), научно-образовательного математического центра (075-02-2020-1483/1) и государственного задания (0729-2020-0055).}
 }

### 2020-08-13 2020/OPTIMA/Answers.tex
Draft for the answers to the reviewers

%% Based on a TeXnicCenter-Template by Gyorgy SZEIDL.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------
%
\documentclass{article}%
%
\usepackage{amsmath}%
\usepackage{amsfonts}%
\usepackage{amssymb}%
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{ulem}
\usepackage{cancel}	
\usepackage[utf8]{inputenc}
\usepackage[english, russian]{babel} % Русские и английские переносы

%-------------------------------------------

\begin{document}

\section*{Answers to the reviewers' comments for the paper ``Parallel global optimization algorithm with uniform convergence for solving a set of constrained global optimization problems''}

We thank very much all the reviewers for their useful comments and suggestions.
Please find below our corrections that we suggest to clarify the issues.

\phantom{f}

\subsection*{Answers to Reviewer 1}

\subsubsection*{Major comments}

\begin{enumerate}
\item 
First of all, this is the Nizhny Novgorod version of the English demonstrated by the authors. ... I recommend that the authors should seek professional help from experienced translators of mathematical texts.

\textbf{Answer}. 
%The paper has been corrected by an English-speaking person.

\item The principal objective of the paper, which is to solve the number (set) of nonconvex problems, is not presented conveniently in the manuscript. There is no mathematical formulation of the problem, say, in the form of an optimization problem, meanwhile the authors tell us about “the optimal way” of solving such a problem, without substantiating this view.

\textbf{Answer}.

\item As well-known now, the solution of (1-3)-dimensional math programming problems, even nonconvex ones, has no considerable difficulties (say, with the help of B\&B methods) and, hence, it is not an attractive value for estimation of the power and effectiveness of algorithms.

\textbf{Answer}.

\item There is no comparison of the computational results with popular approaches in the Global Optimization, such as Random Searches, Genetics Algorithms (for example), and other bioinitiated methods.
Additionally, we recommend that the authors pay attention to the DCA methods (by P.D.Tao and L.Th.An) which are able of producing a stationary (KKT-) vector and even a global solution (sometimes) to DC problems.

\textbf{Answer}.

\end{enumerate}

\subsubsection*{Minor comments}

\begin{enumerate}
\item p. 3. Apparently, $z_k\neq g_s(x_k ) =:z_{ks}$.

\textbf{Answer}.

\item  “To account for the latter” change for “On account of”.

\textbf{Answer}.

\item  p. 4. To put more clear the definition (DF) of $Q_i$ (poor grammar usage).

\textbf{Answer}.

\item  Delete “At that” (???) globally.

\textbf{Answer}.

\item “At the point” put instead of “in the point”.

\textbf{Answer}.

\item  Specify the DF of the trial (what is it?) at the point $x$.

\textbf{Answer}.

\item  Where is a DF of $\nu(x)$?

\textbf{Answer}.

\item  Is it relevant to use at Step 1 “juxtapose”? What does it mean?

\textbf{Answer}.

\item  It seems that in Theorem 1 it is sufficient to assume that $D\neq\emptyset$ (I-st theorem of Weierstrass).

\textbf{Answer}.

\item  Delete “Beginning with some step”, put simply “IF …”

\textbf{Answer}.

\item  Change “admissible” for “feasible”.

\textbf{Answer}.

\item  The equalities (12) are wrong.

\textbf{Answer}.

\item  Delete “where $y_*$ is from (3)”.

\textbf{Answer}.

\item  Completely reformulate Theorem 2.

\textbf{Answer}.

\item  Before Table 1 the numbers (?) $p, S_i, S_t$ are not correctly defined.

\textbf{Answer}.

\item  Delete two last propositions before Table 1.

\textbf{Answer}.

\item  Completely rewrite the conclusion. As it is, it is not written in English.

\textbf{Answer}.

\item  The English and the terminology must be similar to [6] and [16].

\textbf{Answer}.

\end{enumerate}

\subsection*{Answers to Reviewer 2}

\textbf{Comment}. Is it possible to find different solutions by the suggested technique, please, give some comments.

\textbf{Answer}.

\subsection*{Answers to Reviewer 3}

\textbf{Comment}. It's not clear why solving a series of global optimization problems in parallel is an important problem. From my experience such problem statement sounds artificial.

\textbf{Answer}.

\end{document}

### 2020-08-14 2020/OPTIMA/Answers.tex
Some answers in Russian

@@
 
-\section*{Answers to the reviewers' comments for the paper ``Parallel global optimization algorithm with uniform convergence for solving a set of constrained global optimization problems''}
+\section*{Answers to the reviewers' comments 
+%for the paper ``Parallel global optimization algorithm with uniform convergence for solving a set of constrained global optimization problems''
+}
 
 We thank very much all the reviewers for their useful comments and suggestions.
-Please find below our corrections that we suggest to clarify the issues.
+Please find below our answers that we suggest to clarify the issues.
 
-\phantom{f}
+%\phantom{f}
 
@@
 \textbf{Answer}. 
-%The paper has been corrected by an English-speaking person.
+The paper corrected by an English-speaking person.
 
@@
 \textbf{Answer}.
+В рассматриваемых задачах целевая функция и ограничения предполагаются многоэкстремальными и заданными в виде ``черного ящика''.
+Кроме того, подразумевается, что каждое испытание (т.е.
+последовательное вычисление значений ограничений и целевой фукнции) является трудоемкой операцией и может потребовать значительных вычислительных ресурсов.
+Такая постановка часто встречается при решении прикладных задач (например, задач нелинейной аппроксимации; решении систем нелинейных уравнений и неравенств; подборе параметров математических моделей по экспериментальным данным, и т.п.).
+Поэтому задачи липшицевой глобальной оптимизации даже с одной переменной до сих пор вызывают интерес исследователей [1,2], а задачи размерности 2-3 являются стандартными для тестирования и сравнения алгоритмов липшицевой глобальной оптимизации [3].
+
+В статью добавлен комментарий по поводу сложности задач рассматриваемого класса (see ...)
+
+1.
+Calvin, J.M., Chen, Y., {\v Z}ilinskas, A.
+An Adaptive Univariate Global Optimization Algorithm and Its Convergence Rate for Twice Continuously Differentiable Functions.
+J.
+Optim.
+Theory Appl.
+155, 628–636 (2012)
+
+2.
+Sergeyev, Y.D., Mukhametzhanov, M.S., Kvasov, D.E.
+et al.
+Derivative-Free Local Tuning and Local Improvement Techniques Embedded in the Univariate Global Optimization.
+J.
+Optim.
+Theory Appl.
+171, 186--208 (2016)
+
+3. Paulavi{\v c}ius, R., {\v Z}ilinskas, J. Simplicial Global Optimization. Springer Briefs in Optimization. Springer, New York (2014)
 
@@
 \textbf{Answer}.
+Результаты сравнения core Global Search Algorithm с рядом nature-inspired methods при решении нескольких сотен тестовых задач приведены в [1].
+Результаты показывают, что GSA обеспечивает лучшее качество решения задач (решается большее число задач за меньшее число поисковых испытаний).
+В текст статьи добавлен соответствующий комментарий (see page ...).
+
+Одновременно с этим результаты численных экспериментов дополнены сравнением предложенного подхода к решению серий задач оптимизации (на примере многокритериальной задачи оптимизации) и генетического алгоритма для решения MCO problems, реализованного в SciPy (see page ...)
+
+Еще раз отметим, что в работе рассматриваются задачи с функциями вида ``черный ящик'' (i.e., an analytical representation of the functions is not available).
+Поэтому сравнение с методами, использующими аналитическое представление функций задачи (например, основанными на идеях DC-оптимизации, see [2] and references given therein) не будет являться корректным.
+В текст статьи добавлен соответствующий комментарий (see page ...).
+
+1. Sovrasov, V.  Comparison of Several Stochastic and Deterministic Derivative-Free Global Optimization Algorithms.  Lecture Notes in Computer Science. 11548, 70--81 (2019)
+
+2. Pham Dinh T., Le Thi H.A. Recent Advances in DC Programming and DCA. Lecture Notes in Computer Science. 8342, 1--37 (2014)
 

### 2020-08-26 2020/OPTIMA/Answers.tex
Add answers to the majority of the comments

@@
 
-В статью добавлен комментарий по поводу сложности задач рассматриваемого класса (see ...)
+В статью добавлен комментарий по поводу сложности задач рассматриваемого класса (see ...), а также численные эксперименты
+по решению задач размерности 4 (see Section 4.1 and Table 1).
 
@@
 \item p. 3. Apparently, $z_k\neq g_s(x_k ) =:z_{ks}$.
+This formula was reformulated.
 
@@
 \item  “To account for the latter” change for “On account of”.
+The phrase was replaced.
 
@@
 \textbf{Answer}.
+The definition was added to the Section 3. See equation (5).
 
@@
 \textbf{Answer}.
+The definition of the trial was clarified. See the description after the equation (6).
 
@@
 \textbf{Answer}.
+See Section 3, equation (6).
 
@@
 \textbf{Answer}.
+Here “juxtapose” means combine points and the corresponding results of trials into pairs.
+The sentence was clarified.
 
@@
 \textbf{Answer}.
+The original formulation of the theorem was made for the interval [0,1], this assumption is missing indeed.
 
@@
 \textbf{Answer}.
+Deleted.
 
@@
 \textbf{Answer}.
+Done.
 
@@
 \textbf{Answer}.
+The $\inf$ operator is missing here. It was added. See equation (14).
 
@@
 \textbf{Answer}.
+The theorem was reformulated using new aliases for algorithms.
 
@@
 \textbf{Answer}.
+The definitions were added to the Section 4.1.
 

### 2020-08-26 2020/OPTIMA/optima_paper.tex
OPTIMA: Add results of 4d experiments
Add answers to the majority of the comments

@@
   \label{eq:constrained_problem}
-  \varphi(y^*)=\min\{\varphi(y):y\in D \cap \{y: g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}\}.
+  \varphi(y^*)=\min\{\varphi(y):y\in G\},\:G=D \cap \{y: g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}.
 \end{equation}
@@
   \label{eq:many_problems}
-  \min\left\{\varphi_1(y), y\in D_1 \right\}, \min\left\{\varphi_2(y), y\in D_2\right\},...,
-\min\left\{\varphi_q(y), y\in D_q\right\}.
+  \min\left\{\varphi_1(y), y\in G_1 \right\}, \min\left\{\varphi_2(y), y\in G_2\right\},...,
+\min\left\{\varphi_q(y), y\in G_q\right\}.
 \end{equation}
@@
 considered here implies the constructing of a sequence of points
-\(x_k\), which the values of the minimized function or of the constraints \(z_k = g_s(x_k)\) are
-computed in.
-To account for the latter, the index scheme \cite{Strongin2000} was used.
-Let \(Q_0=[0,1]\). The constraint having the index \(j\) is satisfied in all points of the domain
+\(x_k\), which the values of index function \(z_k\) are computed in.
+On account of, the index scheme \cite{Strongin2000} was used.
+Let's assume that each function \(g_i(x), 1 \leqslant i \leqslant m + 1\), is defined and computable
+only in the corresponding subrange \(Q_i \in [a, b]\), where \(\varphi(x)\) is denoted as \(g_{m+1}(x)\) and
+\begin{equation}
+  \label{eq:q}
+  Q_1=[0,1],\: Q_{i+1}=\left\{x\in Q_i:g_{i}(x)\leqslant 0\right\},\:1 \leqslant i \leqslant m, \: Q_{m+2}=\emptyset
+\end{equation}
+In the view of the definition (\ref{eq:q}) the initial problem can be rewritten in the form
 \begin{displaymath}
-  Q_j=\left\{x\in [0,1]:g_j(x)\leq 0\right\},
+  \varphi(x^*)=\min\{g_{m+1}(x):x\in Q_{m+1}\}.
 \end{displaymath}
-which is called feasible for this constraint.
-At that, the feasible domain \(D\) of the initial problem is defined by the equality
- \(D=\cap _{j=0}^{m}Q_{j}\).
-A trial in the point \(x\in [0,1]\) consists in a sequence computing of the values
-\(g_{1}(x),...,g_{\nu }(x)\)
-where the values of the indices \(\nu\) are defined by the conditions:
-\(x\in Q_{j},0\leqslant j<\nu ,x\notin Q_{\nu }\).
-The first observed violation of any constraint terminates the trial at the point \(x\).
-In the case when the point \(x\) is feasible, i.e. \(x\in D\), the trial includes the computing of all
-functions of the problem.
-At that, the value of the index is accepted as \(\nu =m+1\).
-The pair \(\nu =\nu (x), z=g_{\nu }(x)\) where the index \(\nu\) falls into the borders \(1\leqslant
-\nu \leqslant m+1\) is called the result of the trial at the point \(x\).
+
+The index scheme assigns a number called index \(\nu(x)\) to the each point of the search sequence \(x_k\):
+\begin{equation}
+  \nu(x) = i:x\in Q_i, x\not\in Q_{i+1},\:1\leqslant i\leqslant m + 1.
+\end{equation}
+
+In order to obtain an index of the point \(x\) we have to perform a trial defined via the following steps:
+\begin{enumerate}
+  \item Consequently compute functions \(g_i(x),1 \leqslant i \leqslant m\) until for some \(i^*\) \(g_{i^*}(x)>0\) or until \(i=m+1\).
+  Set \(\nu(x)=i^*\) or \(\nu(x)=m+1\) if \(g_i(x) \leqslant 0, 1 \leqslant i \leqslant m\).
+  Therefore, \(\nu(x)\) is the number of the first violated constraint at the point \(x\) or \(m+1\) if
+  all constraints are satisfied at \(x\).
+  \item Return pair \(\nu(x), z=g_{\nu(x)}(x)\) as a result of trial.
+\end{enumerate}
 
@@
 Here \(M=\max_{}^{}\left\{\nu (x):x\in [0,1]\right\}\) and \(g_{M}^{*}=\min
-_{}^{}\left\{g_{M}(x):x\in \cap _{i=0}^{M-1}Q_{i}\right\}\).
+_{}^{}\left\{g_{M}(x):x\in Q_{M}\right\}\).
 Because of the definition of the number \(M\), the problem of finding \(g_{M}^{*}\) always
 has a solution. If \(M=m+1\), then \(g_{M}^{*}=\varphi(x^{*})\).
-The function \(\psi (x)\) satisfies the H\"{o}lder condition within the sets
-\(\cap _{i=0}^{j}Q_{i},0\leq j\leq M-1\)
-with the constant 1, and \(\psi (x)\) can have jump discontinuities at the boundaries
-of these sets.
+The function \(\psi (x)\) satisfies the H\"{o}lder condition on the set
+\(Q_1\) with the constant 1, and \(\psi (x)\) can have jump discontinuities at the boundaries
+of the sets \(Q_i,\:1\leqslant i \leqslant m+1\).
 Despite the values of the H\"{o}lder constants \(H_k\) and the value \(g_{M}^{*}\) are not
@@
 \end{equation}
-and juxtapose the values \(z_{i}=g_{\nu }(x_{i}),\nu =\nu (x_{i}),i={\overline {1,k}}\) to these
-ones.
+and juxtapose to them the values \(z_{i}=g_{\nu }(x_{i}),\nu =\nu (x_{i}),i={\overline {1,k}}\) computed at these points.
 
@@
 viewpoint of the characteristics (\ref{step3_1}) that allows distributing the resources of the
-method among the problems dynamically.
+method among the problems dynamically. Let's denote this method as MIAGS.
 
@@
   \begin{enumerate}
-    \item The problem (\ref{eq:constrained_problem}) has a solution.
+    \item \(D\ne\emptyset\), the problem (\ref{eq:constrained_problem}) has a solution.
     \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
 respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
-    \item Beginning with some step (i.e. if \(k\) from (\ref{eq:points}) is sufficiently large),
+    \item If \(k\) from (\ref{eq:points}) is sufficiently large,
     the values \(\mu_\nu\) from (\ref{step2}) satisfy the inequalities
@@
   by the index method while solving the problerm
-  (\ref{eq:constrained_problem}) is admissible and satisfies the conditions
+  (\ref{eq:constrained_problem}) is feasible and satisfies the conditions
 \begin{equation}
   \label{eq:conv_cond}
-  \varphi(\overline{y})=\{ \varphi(y^k): g_i(y^k)\leqslant 0,1\leqslant j\leqslant m, k=1,2,\dots\}=\varphi(y^*),
+  \varphi(\overline{y})=\inf\{ \varphi(y^k): g_i(y^k)\leqslant 0,1\leqslant i\leqslant m, k=1,2,\dots\}=\varphi(y^*).
 \end{equation}
-  where \(y^*\) is from (\ref{eq:constrained_problem}).
 \end{theorem}
@@
 
-\begin{theorem} (On convergence conditions) Let the conditions of the Theorem \ref{th:single_conv}
-  are satisfied for all of \(q\) problems from the set (\ref{eq:many_problems}) separately.
-  Then solving all of the \(q\) problems by the algorithm from Section \ref{sec:method} will
-  generate \(q\) infinite sequences, for the limit points of which the Theorem \ref{th:single_conv}
-  will be also true.
+\begin{theorem} (On convergence conditions of MIAGS) Let for each problem \(i,\:1\leqslant i\leqslant q\) from (\ref{eq:many_problems})
+  the conditions 1-3 of the Theorem \ref{th:single_conv} are true, i.e. each problem can be solved by IAGS.
+  Then solving all of the \(q\) problems by MIAGS will
+  generate \(q\) infinite sequences \(\{y^k_i\},\:1\leqslant i\leqslant q\), such that
+  \begin{displaymath}
+    \varphi_i(\overline{y_i})=\inf\{ \varphi(y^k_i): g^i_j(y^k_i)\leqslant 0,1\leqslant j\leqslant m_i, k=1,2,\dots\}=\varphi_i(y^*_i).
+  \end{displaymath}
 \end{theorem}
@@
 with predefined number of global optimums.
-In combination with GCGen, two sets of 100 problems each of the dimensionalities 2 and 3
-were generated.
-Each problem had two constraints.
-Also, in order to demonstrate the properties of the method to be preserved at essentially
-different properties of the problems, a mixed class consisting of 50 problems with
-two-dimensional GKLS functions and 50 problems with \(F_{GR}\) ones was generated.
+In combination with GCGen, the following sets of problems were generated:
+\begin{itemize}
+  \item 100 2-dimensional problems with two constraints;
+  \item 100 3-dimensional problems with two constraints
+  \item 20 4-dimensional problems with two constraints;
+  \item a mixed class consisting of 50 problems with
+  two-dimensional GKLS functions and 50 problems with \(F_{GR}\).
+  This set is generated in order to demonstrate the properties of the method to be preserved at essentially
+  different properties of the problems.
+\end{itemize}
+
 Examples of contour plots of the considered problems are presented in Fig. \ref{fig:isolines}.
@@
 
-The results of solving the test problems by the sequential and parallel versions of the modified
-IAGS for solving the sets of problems are presented in Table~\ref{tab:speedup}.
+The results of solving the test problems by the sequential and parallel versions of the
+MIAGS are presented in Table~\ref{tab:speedup}.
 For all two-dimensional classes of problems, the parameter \(r=4.7\).
-In the case of the three-dimensional problems, \(r=4.7\) and also to improve the convergence speed
+In the case of the three and four-dimensional problems, \(r=4.7\) and also to improve the convergence speed
 \(\varepsilon\)-reservation technique from \cite{Strongin2000} Chapter 8.3 with
@@
 
-One can see from Table \ref{tab:speedup} that the speedup in the iterations \(S_i\) increased
-linearly with increasing number of threads \(p\) whereas the speedup in time \(S_t\) increased
+One can see from Table \ref{tab:speedup} that the speedup in the iterations \(S_i=\frac{iters(p=1)}{iters(p=i)}\) increased
+linearly with increasing number of threads \(p\) whereas the speedup in time \(S_t=\frac{time(p=1)}{time(p=i)}\) increased
 not so fast that points to a non-ideal implementation of the algorithm.
@@
     \hline
+    GKLS based 4d \
+      & 1 & 14021720 & 15806.6 & -    & - \\
+      & 2 & 6313070 & 7254.85  & 2.22 & 2.18 \\
+      & 4 & 3479344 & 4932.55  & 4.03 & 3.20 \\
+      & 6 & 2783339 & 3955.38  & 5.04 & 3.99 \\
+    \hline
   \end{tabular}
@@
 
-In order to demonstrate the uniform convergence, all test problems have been solved by IAGS
-in the regime of solving separate problems as well.
+In order to demonstrate the uniform convergence of MIAGS, all test problems have been solved by IAGS as well.
 Fig. \ref{fig:devs_mixed} presents the plots of mean and maximum distances from the actual
@@
 two different generators separately (solid line) and together (dashed line).
-In spite of an essential differences in the structure of problems, the modified IAGS decreased
+In spite of an essential differences in the structure of problems, the MIAGS decreased
 the maximum deviations of the estimates from the optima as well as the mean ones much faster.
@@
 the computational resources were distributed among the problems uniformly;
-  \item all problems are solved simultaneously using generalized IAGS with a preset limit of
+  \item all problems are solved simultaneously using MIAGS with a preset limit of
 2500 trials.
@@
     \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{pic/single_mco.pdf}}}
-    \subfloat[IAGS for the set of problems]{
+    \subfloat[MIAGS for the set of problems]{
     \label{fig:mco_pareto_2} {\includegraphics[width=.5\textwidth]{pic/multi_mco.pdf}}}

### 2020-09-15 2020/OPTIMA/Answers.tex
OPTIMA: clarify definition of uniform convergence

@@
 
-Одновременно с этим результаты численных экспериментов дополнены сравнением предложенного подхода к решению серий задач оптимизации (на примере многокритериальной задачи оптимизации) и генетического алгоритма для решения MCO problems, реализованного в SciPy (see page ...)
-
 Еще раз отметим, что в работе рассматриваются задачи с функциями вида ``черный ящик'' (i.e., an analytical representation of the functions is not available).
@@
 \item p. 3. Apparently, $z_k\neq g_s(x_k ) =:z_{ks}$.
+\textbf{Answer}.
 This formula was reformulated.
 
+\item  “To account for the latter” change for “On account of”.
 \textbf{Answer}.
-
-\item  “To account for the latter” change for “On account of”.
 The phrase was replaced.
 
-\textbf{Answer}.
-
 \item  p. 4. To put more clear the definition (DF) of $Q_i$ (poor grammar usage).
-
 \textbf{Answer}.
@@
 \item  Delete “At that” (???) globally.
-
 \textbf{Answer}.
+Deleted.
 
@@
 \textbf{Answer}.
+Fixed.
 
@@
 \textbf{Answer}.
+Conclusion was rewritten.
 
@@
 \textbf{Answer}.
+Definitions and terminology in the Section 3 was corrected.
 

### 2020-09-15 2020/OPTIMA/optima_paper.tex
OPTIMA: clarify definition of uniform convergence

@@
 \end{equation}
+
+As \cite{BarkalovStrongin2018}, this work aims to formulate a method that will ensure a uniform convergence to the
+solutions of all problems in the series. Here and below uniform convergence is taken to
+mean proportional decreasing of the distance between the best iteration point and the
+global optimizer in all the problems in the series:
+\begin{equation}
+  \exists \varepsilon > 0: \forall s>1, \forall i,j\in\{1,\dots,q\}\;
+    \frac{\Vert \tilde{y_i}(s)^* - y_i*\Vert_\infty}{\Vert \tilde{y_j}(s)^* - y^*_j\Vert_	\infty} \leqslant \varepsilon,
+\end{equation}
+where \(s\) is the number of step of an optimization method, \(\tilde{y_i}(s)^*\) is the best iteration point
+in problem \(i\) from the set (\ref{eq:many_problems}) at the step \(s\).
 

### 2020-09-16 2020/OPTIMA/Answers.tex
OPTIMA: update answers

@@
 \textbf{Answer}.
+Clarification of the objective of this work was added to the end of the Section 2.
 
@@
 Результаты показывают, что GSA обеспечивает лучшее качество решения задач (решается большее число задач за меньшее число поисковых испытаний).
-В текст статьи добавлен соответствующий комментарий (see page ...).
+В текст статьи добавлен соответствующий комментарий (see page 11).
 
@@
 Поэтому сравнение с методами, использующими аналитическое представление функций задачи (например, основанными на идеях DC-оптимизации, see [2] and references given therein) не будет являться корректным.
-В текст статьи добавлен соответствующий комментарий (see page ...).
+В текст статьи добавлен соответствующий комментарий (see page 2).
 
@@
 \textbf{Answer}.
+The presented method converges to all the global optimal solutions at once, but numerically it's hard
+to obtain them all with a guarentee. That happens because of luck of an appropriate stopping criterion.
 
@@
 
+A serie of global optimization problems can arise from an mixed-integer problem or
+after scalarization [1] of an multi-objective problem [2]. In both cases it's needed to
+solve all the problems from the serie approximetely with the same quality, because all of
+them are equally important. To speedup this process under constrained computational resources,
+we can apply parellelism in couple with a resource balancing approach to guarantee equal
+quality of solutions in all the problems by the end of the optimization process.
+
+1. Barkalov, K. and I. Lebedev. “Parallel Global Optimization for Non-convex Mixed-Integer Problems.” RuSCDays (2019).
+
+2.
+Jahn J.
+Scalarization in Multi Objective Optimization.
+In: Serafini P.
+(eds) Mathematics of Multi Objective Optimization.
+International Centre for Mechanical Sciences (Courses and Lectures), vol 289.
+Springer, Vienna (1985).
+
 \end{document}

### 2020-09-16 2020/OPTIMA/optima_paper.tex
OPTIMA: update answers

@@
 whereas to \textit{guarantee} the finding of the global optimum the optimization
-methods have to accumulate the information on the behaviour of the objective function in the
+methods have to accumulate the information on the behavior of the objective function in the
 whole search domain \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}.
+In some cases, like for DC or QP problems, one can propose more effective methods and even
+analytical optimality conditions \cite{PhamDinh2014, qpBook}, but when we have to deal with black box functions
+(i.e., an analytical representation of the functions is not available) these methods are barely applicable.
+
 The solving of a series of such problems at limited computational resources is even more
@@
 In order to demonstrate the uniform convergence of MIAGS, all test problems have been solved by IAGS as well.
+IAGS is comparable with other stochastic and deterministic derivative-free global optimization algorithms \cite{Sovrasov2019}. Thus it
+provides a strong baseline in solving a single global optimization problem.
 Fig. \ref{fig:devs_mixed} presents the plots of mean and maximum distances from the actual

### 2020-09-25 2020/RSD_GPU/RSD2020.tex
fix comments from program commettee

@@
 %
-\title{Adaptive global optimization using graphics accelerators\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
+\title{Adaptive Global Optimization Using Graphics Accelerators
+%\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}
 }
 %
-\titlerunning{Adaptive global optimization}  % abbreviated title (for running head)
+\titlerunning{Adaptive Global Optimization}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 \author{
-Konstantin A. Barkalov%(ORCID 0000-0001-5273-2471)
+Konstantin Barkalov%(ORCID 0000-0001-5273-2471)
 $^1$ \and
-Ilya G. Lebedev \letter%(ORCID 0000-0002-8736-0652)
+Ilya Lebedev \letter%(ORCID 0000-0002-8736-0652)
 $^1$ \and 
-Vassili V. Toropov%(ORCID 0000-0002-7017-5055)
+Vassili Toropov%(ORCID 0000-0002-7017-5055)
 $^{1,2}$
@@
 %
-\authorrunning{K.A. Barkalov \and I.G. Lebedev \and V.V. Toropov} % abbreviated author list (for running head)
+\authorrunning{K. Barkalov \and I. Lebedev \and V. Toropov} % abbreviated author list (for running head)
 %
@@
 
-\section{Multidimensional parallel global search algorithm}\label{Sec_GSA}
+\section{Multidimensional Parallel Global Search Algorithm}\label{Sec_GSA}
 Let's consider the problem of finding the global minimum of the \(N\)-dimensional function \(\varphi(y)\) in the hyperinterval  \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
@@
 
-\section{Dimensionality reduction schemes in global optimization problems}
+\section{Dimensionality Reduction Schemes in Global Optimization Problems}
 
-\subsection{Dimensionality reduction using multiple mappings}
+\subsection{Dimensionality Reduction Using Multiple Mappings}
 
@@
 
-\subsection{Recursive dimensionality reduction scheme}
+\subsection{Recursive Dimensionality Reduction Scheme}
 
@@
 
-\subsection{Adaptive dimensionality reduction scheme}
+\subsection{Adaptive Dimensionality Reduction Scheme}
 
@@
 
-\section{GPU implementation}
+\section{GPU Implementation}
 
-\subsection{General scheme}
+\subsection{General Scheme}
 
@@
 
-\subsection{Organization of parallel computing}
+\subsection{Organization of Parallel Computing}
 
@@
 
-\section{Numerical experiments}
+\section{Numerical Experiments}
 
@@
 
+\medskip
+
+\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,16-11-10150.
+
 %

### 2020-09-27 2020/OPTIMA/slides/slides.tex
OPTIMA: Add a stub for slides

--- 2020-05-28 2020/PAVT_Sovrasov/slides/slides.tex
@@
 
-\graphicspath{ {../images/}{img/} }
+\graphicspath{ {../pic/}{img/} }
 
@@
 
-\title{Параллельный алгоритм для получения равномерного приближения решений множества задач глобальной оптимизации с нелинейными ограничениями}
-\author{\underline{\textbf{В.В.~Соврасов}} \and \textbf{К.А.~Баркалов}}
-\institute{Нижегородский государственный университет им. Н.И. Лобачевского}
+\title{Parallel global optimization algorithm with uniform convergence
+for solving a set of constrained global optimization problems}
+\author{\underline{\textbf{V.V.~Sovrasov}} \and \textbf{K.A.~Barkalov}}
+\institute{Nizhny Novgorod State University}
 \date{}
@@
 \begin{frame}
-  \frametitle{Постановка задачи}
+  \frametitle{Problem statement}
   \begin{columns}
@@
       \end{displaymath}
-      \(\varphi(y),\:g_j(y)\) -- многоэкстремальные функции, удовлятворяющие условию Липшица:
+      \(\varphi(y),\:g_j(y)\) -- multi-extremal functions, satisfying the Lipschitz condition:
       \begin{displaymath}
@@
       \end{displaymath}
-      где \(L>0\) константа Липшица и \(||\cdot||\) обозначает \(l_2\) норму в пространстве \(\mathbb{R}^N\).
+      where \(L>0\) is the Lipschitz constant and \(||\cdot||\) denotes the \(l_2\)-norm in \(\mathbb{R}^N\).
     \end{column}
@@
 \begin{frame}
-  \frametitle{Постановка задачи}
+  \frametitle{Problem statement}
   Далее будем интересоваться решением серии из \(q\) задач глобальной оптимизации с нелинейными ограничениями:
@@
 \begin{frame}
-  \frametitle{Редукция размерности}
+  \frametitle{Dimension reduction}
   \begin{center}

### 2020-09-28 2020/OPTIMA/slides/slides.tex
OPTIMA: update slides

@@
   \frametitle{Problem statement}
-  Далее будем интересоваться решением серии из \(q\) задач глобальной оптимизации с нелинейными ограничениями:
+  Let's consider a set of \(q\) global optimization problems with nonlinear constraints:
   \begin{displaymath}
@@
   \end{displaymath}
-  Подобные серии задач могут возникнуть, например, в следующих случаях:
+  For an instance, such sets of problems can arise in the following cases:
   \begin{itemize}
-    \item задача глобальной оптимизации с дискретным параметром;
-    \item  решение задачи многокритериальной оптимизации методом свертки критериев.
+    \item continuous global optimization problem with a discrete parameter which takes a narrow range of values;
+    \item a set of problems, obtained after scalarization of a multi-objective problem.
   \end{itemize}
-\textbf{Возможные методы решения}:
+\textbf{Possible approaches}:
   \begin{itemize}
-    \item решать каждую задачу независимо;
-    \item разработать метод оптимизации, который решает все задачи из множества в совокупности, в каждый
-    момент времени приоретизируюя одну из них.
+    \item solve each problem independently;
+    \item develop an optimiaztion method, which solves all the problems "simultaneously", at each moment of time
+    prioritizing one of the problems.
   \end{itemize}
@@
   \begin{center}
-  Кривая Пеано \(y(x)\) позволяет уменьшить размерность многомерного пространства до 1:
+  Peano-type curve \(y(x)\) allows to reduce the dimension of the original problem:
   \begin{gather}
@@
   \end{gather}
-  \(y(x)\) является негладкой кривой, отображающей отрезок \([0,1]\) на гиперкуб \(D_e\).
+  \(y(x)\) is non-smooth function which continuously maps the segment \([0,1]\) to the hypercube \(D_e\).
   \begin{figure}[ht]
@@
 \begin{frame}
-  \frametitle{Алгоритм глобальной оптимизации}
-  Метод оптимизации генерирует последовательность точек \(\{x_k:x_k\in[a,b]\}\) и состоит в выполнении следующих шагов:
+  \frametitle{Univariate Algorithm of Global Search}
+  Optimization method generates search sequence \(\{x_k:x_k\in[a,b]\}\) and consists of the following steps:
   \begin{enumerate}
     \setlength{\itemindent}{.1in}
-    \item[Шаг 1.] Упорядочить поисковую информацию (одномерные точки) по возрастанию.
-    \item[Шаг 2.] Для каждого интервала \((x_{i-1}, x_i)\) вычислить величину \(R(i)\), называемую характеристикой.
-    \item[Шаг 3.] Выбрать интервал \((x_{t-1}, x_{t})\) с наибольшей характеристикой и
-    провести испытание (вычислить органичения и целевую функцию) в точке \(x^{k+1}\), выбранной с помощью решающего правила \(d\):
+    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
+    \item[Step 2.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
+    \item[Step 3.] Choose an interval with the greatest characteristic, \((x_{t-1}, x_{t})\), and
+    make a trial (compute the constraints and objective) at the point \(x^{k+1}\), chosen according with decision rule \(d\):
     \begin{displaymath}
@@
     \end{displaymath}
-    \item[Шаг 4.] Если \(x_{t}-x_{t-1}<\varepsilon\), остановить метод.
+    \item[Step 4.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
   \end{enumerate}
-  \textit{\footnotesize	{Детальное описание метода: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
+  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Алгоритм, решающий множество задач}
-  Модификация метода для решения множества задач:
+  \frametitle{Algorithm for solving a set of problems}
+  The method relies on the previously introduces AGS and consists of the following steps:
   \begin{itemize}
-    \item Создать \(q\) копий характеристического АГО.
-    \item Использовать \(q\) синхронно работающих копий АГО с тем лишь отличием, что на шаге 3 при выборе
-    интервала с наилучшей характеристикой, выбор будет осуществляться из всех интервалов, которые
-    породили на данный момент \(q\) копий АГО.\
-    \item Если наибольшая характеристика соответствует
-    задаче \(i\), то выполняется шаг 3 в копии метода с номером \(i\), а остальные копии метода простаивают.
-    \item Для реализации параллельности в предыдущем пункте можно выбрать \(p\) характеристик и провести
-    испытания в \(p\) задачах одновременно.
+    \item Create \(q\) copies of AGS.
+    \item Run \(q\) synchronously working copies of AGS, but instead of performing Step 3, stop all
+    the methods and select an interval with the highest characteristic among intervals from all the methods.
+    \item If the interval with the highest characteristic belongs to problem i \(i\), execute Step 3 in the \(i^{th}\) copy of AGS.
+    Other copies skip Step 3.
+    \item To implement parallel processing, method picks \(p\) intervals at the previous two steps
+    and performs \(p\) trials simultaneously.
   \end{itemize}
-  Чтобы приведённая схема работала, требуется \textbf{сравнимость характеристик интервалов в разных задачах}.
+  The described method relies to \textbf{comparability of characteristics in different problems}.
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Индексная схема учёта ограничений}
-  Вместо исходной задачи рассмотрим следующую задачу без функциональных ограничений:
+  \frametitle{Index method of handling inequality constraints}
+  Instead of the original problems with functional constraints the index scheme considers the following unconstrained problem:
   \begin{displaymath}
@@
       \psi (x^{*})=\min_{x\in [0;1]}\psi (x), \\
-      \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu }&\nu <M\\(g_{M}(x)-g_{M}^{*})/H_{M}&\nu =M\end{cases}}
+      \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu }&\nu <M\\(g_{M}(x)-g_{M}^{*})/H_{M}&\nu =M\end{cases}},
     \end{array}
   \end{displaymath}
-  где при \(\nu=m+1\;g_\nu(x)=\varphi(x)\).
+  where if \(\nu=m+1\;,g_\nu(x)=\varphi(x)\).
 
-  При использовании индексной схемы характеристики определяются следующим образом:
+  Characteristics are defined as follows:
   \begin{displaymath}
@@
   \end{displaymath}
-  где \(z_i=\psi(x_i)\). Можно показать, что при использовании таких характеристик метод будет выбирать интервалы из всех \(q\) задач, тем самым решая их все, а не только некоторое подмножество задач.
+  where \(z_i=\psi(x_i)\).
+\end{frame}
+
+\begin{frame}
+  \frametitle{Index method of handling inequality constraints}
 \end{frame}

### 2020-09-29 2020/OPTIMA/slides/slides.tex
OPTIMA: fixes, update slides

@@
         \begin{array}{cr}\\
-          \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
-          D=\{y\in \mathbb{R}^N:a_i\leqslant y_i\leqslant{b_i}, \\ 1\leqslant{i}\leqslant{N},\:g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}
+          \varphi(y^*)=\min\{\varphi(y):y\in G\},\:G=D \cap \{y:\\
+           g_j(y)\leqslant 0, 1\leqslant j\leqslant m\},
         \end{array}
       \end{displaymath}
+      where \(D\) is a hyperrectangle,
       \(\varphi(y),\:g_j(y)\) -- multi-extremal functions, satisfying the Lipschitz condition:
@@
   Let's consider a set of \(q\) global optimization problems with nonlinear constraints:
-  \begin{displaymath}
-    \min\left\{\varphi_1(y), y\in D_1 \right\}, \min\left\{\varphi_2(y), y\in D_2\right\},..., \min\left\{\varphi_q(y), y\in D_q\right\}.
-  \end{displaymath}
-  For an instance, such sets of problems can arise in the following cases:
+  \begin{equation}
+    \label{eq:many_problems}
+    \min\left\{\varphi_1(y), y\in G_1 \right\}, \min\left\{\varphi_2(y), y\in G_2\right\},...,
+  \min\left\{\varphi_q(y), y\in G_q\right\}.
+  \end{equation}
+  This work aims to formulate a method that will ensure the uniform convergence to the
+  solutions of all problems in the series:
+  \begin{equation}
+    \exists \varepsilon > 0: \forall s>1, \forall i,j\in\{1,\dots,q\}\;
+      \frac{\Vert \overline{y_i}^*(s) - y_i^*\Vert_\infty}{\Vert \overline{y_j}^*(s) - y_j^*\Vert_	\infty} \leqslant \varepsilon,
+  \end{equation}
+  where \(s\) is the number of step of an optimization method, \(\overline{y_i}^*(s)\) is the best iteration point
+  in problem \(i\) from the set (\ref{eq:many_problems}) at the step \(s\).
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Problem statement}
+  Such sets of problems can arise in the following cases:
   \begin{itemize}
@@
   \frametitle{Index method of handling inequality constraints}
+  \begin{enumerate}
+    \item \(D\ne\emptyset\), the problem (\ref{eq:constrained_problem}) has a solution.
+    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
+respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
+    \item If \(k\) from (\ref{eq:points}) is sufficiently large,
+    the values \(\mu_\nu\) from (\ref{step2}) satisfy the inequalities
+    \begin{equation}
+      r_\nu\mu_\nu > 2^{3-1/N}L_\nu \sqrt{N+3},\: 1\leqslant \nu \leqslant m + 1.
+    \end{equation}
+  \end{enumerate}
+  \begin{theorem} (On convergence conditions of MIAGS) Let for each problem \(i,\:1\leqslant i\leqslant q\) from (\ref{eq:many_problems})
+    the conditions 1-3 of the Theorem \ref{th:single_conv} are true, i.e. each problem can be solved by IAGS.
+    Then solving all of the \(q\) problems by MIAGS will
+    generate \(q\) infinite sequences \(\{y^k_i\},\:1\leqslant i\leqslant q\), such that
+    \begin{displaymath}
+      \varphi_i(\overline{y_i})=\inf\{ \varphi(y^k_i): g^i_j(y^k_i)\leqslant 0,1\leqslant j\leqslant m_i, k=1,2,\dots\}=\varphi_i(y^*_i).
+    \end{displaymath}
+  \end{theorem}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Пример решения многокритериальной задачи}
-  Рассматриваемая задача:
+  \frametitle{An example of solving a multi-objective problem}
+  Consider the following problem:
   \begin{displaymath}
@@
   \end{displaymath}
-  После использования свёртки Гермейера для скаляризации задача примет вид:
+  After scalarization the problem has the following form:
   \begin{displaymath}
@@
 \end{displaymath}
-Для численного построения множества Парето выберем
-100 наборов коэффициентов \((\lambda_1,\lambda_2)\) таких, что
+To build the Pareto set numerically, consider
+100 pairs of the variables \((\lambda_1,\lambda_2)\) such that
 \(\lambda_1^i=i h,\: \lambda_2^i=1-\lambda_1^i,\: h=10^{-2},i=\overline{1, 100}\).
@@
 \begin{frame}
-  \frametitle{Пример решения многокритериальной задачи}
+  \frametitle{An example of solving a multi-objective problem}
   \vspace*{-1.0cm}
@@
       \vspace*{-0.5cm}
-      \subfloat[Раздельное решение задач, \(SP_{single}=0.984\)]{
+      \subfloat[IAGS, \(SP_{single}=0.984\)]{
       \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{single_mco.pdf}}}
-      \subfloat[Решение множества задач, \(SP_{multi}=0.749\)]{
+      \subfloat[MIAGS, \(SP_{multi}=0.749\)]{
       \label{fig:mco_pareto_2} {\includegraphics[width=.5\textwidth]{multi_mco.pdf}}}
-      \caption{Численные оценки множества Парето, полученные после 2500 испытаний}
+      \caption{Numerical estimations of the Pareto set, obtained after 2500 trails}
       \label{fig:mco_pareto}
@@
 \begin{frame}
-  \frametitle{Тестовые многомерные задачи с ограничениями}
+  \frametitle{Generated test problems}
   \begin{columns}
     \begin{column}{0.5\textwidth}
-      Тестовые наборы были получены с помощью системы GCGen, позволяющей
-      сторить многомерные задачи с ограничениями из заданных функций, и нескольких
-      известных генераторов тестовых задач (GKLS, \(F_{GR}\)).
-      Характеристики каждого из сгенерированных наборов задач:
+      Sets of test problems are obtained by applying the GCGen system to
+      functions produced by generators of unconstrained test problems.
+      The main features of the generated sets:
       \begin{itemize}
         \item \(q\)=100;
-        \item Размерность 2 или 3;
-        \item Базовые функции только GKLS или \(F_{GR}\), или их комбинация;
-        \item Количество ограничений: 2.
+        \item Dimension 2, 3 or 4;
+        \item Base fuctions GKLS or \(F_{GR}\), or a combination of them;
+        \item Number of nonlinear constraints: 2.
       \end{itemize}
@@
   \end{columns}
-  \footnotesize{Система GCGen доступна по ссылке \textit{https://github.com/UNN-ITMM-Software/GCGen}}
+  \footnotesize{GCGen is available here \textit{https://github.com/UNN-ITMM-Software/GCGen}}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Программное и техническое обеспечение}
+  \frametitle{Hardware and software environment}
   \begin{center}
-    Реализация параллельного метода была выполнена на языке C++ с использованием технологии OpenMP
-    для распареллеливания процесса проведения испытаний на общей памяти.
+    The parallel method was implemented on C++ language with the use of OpenMP
+    technology for organizing parallelism on shared memory.
 
-    Все вычислительные
-    эксперименты проведены на машине со следующей конфигурацией: Intel Core i7-7800X (6 cores) CPU, 64GB RAM, Unubtu 16.04 ОS, GCC 5.5 compiler.
+    All the numerical experiments were carried out on a computer with the following
+    configuration: Intel Core i7-7800X (6 cores) CPU, 64GB RAM, Unubtu 16.04 ОS, GCC 5.5 compiler.
   \end{center}
@@
 \begin{frame}
-  \frametitle{Результаты экспериментов на наборах синтетических задач}
+  \frametitle{Result of solving sets of synthetic problems}
   \begin{figure}[ht]
@@
       \label{fig:avg_dev} {\includegraphics[width=.5\textwidth]{mixed_2d_avg.pdf}}}
-      \caption{Динамика величин средней по задачам и максимальной точности в процессе решения множества двухмерных задач,
-      порождённых двумя разными генераторами GKLS и \(F_{GR}\)}
+      \caption{Dynamic of the average and maximal errors, during solving of problems
+      generated on different types of base functions: GKLS и \(F_{GR}\)}
       \label{fig:devs_mixed}
@@
 \begin{frame}
-  \frametitle{Результаты экспериментов на наборах синтетических задач}
+  \frametitle{Result of solving sets of synthetic problems}
   \begin{table}
@@
       %\cline{1-8}\noalign{\smallskip}
-      Класс задач & \textit{p} & Количество итераций & Время, с & \(S_i\) & \(S_t\)   \\
+      Problems class & \textit{p} & Number of iterations & Time, s & \(S_i\) & \(S_t\)   \\
       %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
@@
       \hline
+    \end{tabular}
+  \end{table}
+\end{frame}
+\begin{frame}
+  \frametitle{Result of solving sets of synthetic problems}
+  \begin{table}
+    \centering
+    \begin{tabular}{c|c|c|c|c|c}
+      %\cline{1-8}\noalign{\smallskip}
+      Problems class & \textit{p} & Number of iterations & Time, s & \(S_i\) & \(S_t\)   \\
+      %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
+      \hline
       GKLS-based 3d \
@@
       \hline
+      GKLS-based 4d \
+        & 1 & 14021720 & 15806.6 & -    & - \\
+        & 2 & 6313070 & 7254.85  & 2.22 & 2.18 \\
+        & 4 & 3479344 & 4932.55  & 4.03 & 3.20 \\
+        & 6 & 2783339 & 3955.38  & 5.04 & 3.99 \\
+      \hline
     \end{tabular}
@@
 \end{frame}
-
 \begin{frame}
-  \frametitle{Заключение}
-  В ходе работы были достигнуты следующие результаты:
+  \frametitle{Conclusions}
     \begin{itemize}
-      \item реализована поддержка нелинейных ограничений в алгоритме, решающeм
-      множество задач глобальной оптимизации;
-      \item проведены численные эксперименты, демонстрирующие преимущество рассматриваемого подхода в скорости сходимости на всём множестве задач в среднем над решением задач по отдельности;
-      \item показана эффективность совместного решения множества задач на примере решения многокритериальной задачи с нелинейными ограничениями.
-    \end{itemize}
-
-    В ходе дальнейшей работы планируется:
-    \begin{itemize}
-      \item улучшить текущую реализацию алгоритма, сократив расходы на содержание поисковой информации для множества задач и тем самым улучшив
-      показатели параллельного ускорения по времени;
-      \item реализовать версию рассматриваемого алгоритма, работающего на распределенной памяти.
+      \item implemented support of nonlinear functional constraints in the method
+      for solving a set of global optimization problems;
+      \item convergence of the implemented method was theoretically proven;
+      \item conducted numerical experiments on sets, demonstrating uniform convergence (in the mentioned sense) and
+      parallel speedup;
+      \item demonstrated efficiency of the implemented method in case of solving a multi-objective problem.
     \end{itemize}
@@
   \begin{center}
-    \Large{Контакты:}
+    %\Large{Контакты:}
 \vspace{0.5cm}

### 2020-10-01 2020/OPTIMA/slides/slides.tex
OPTIMA: final update of slides

@@
       \end{displaymath}
-      where \(D\) is a hyperrectangle,
+      where \(D\) is a hyperrectangle in \(\mathbb{R}^N\),
       \(\varphi(y),\:g_j(y)\) -- multi-extremal functions, satisfying the Lipschitz condition:
@@
   \begin{itemize}
-    \item continuous global optimization problem with a discrete parameter which takes a narrow range of values;
+    \item continuous global optimization problem with a discrete parameter, which takes a narrow range of values;
     \item a set of problems, obtained after scalarization of a multi-objective problem.
@@
 \begin{frame}
+  \frametitle{Univariate Algorithm of Global Search}
+  Sufficient convergence conditions:
+  \begin{enumerate}
+    \item \(G\ne\emptyset\), the source problem has a solution.
+    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
+respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
+    \item During the optimization process estimations of \(L_i\) becomes large enough.
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}
   \frametitle{Algorithm for solving a set of problems}
-  The method relies on the previously introduces AGS and consists of the following steps:
+  The method relies on the previously introduced AGS and consists of the following steps:
   \begin{itemize}
@@
     the methods and select an interval with the highest characteristic among intervals from all the methods.
-    \item If the interval with the highest characteristic belongs to problem i \(i\), execute Step 3 in the \(i^{th}\) copy of AGS.
+    \item If the interval with the highest characteristic belongs to problem \(i\), execute Step 3 in the \(i^{th}\) copy of AGS.
     Other copies skip Step 3.
@@
   where \(z_i=\psi(x_i)\).
-\end{frame}
-
-\begin{frame}
-  \frametitle{Index method of handling inequality constraints}
-  \begin{enumerate}
-    \item \(D\ne\emptyset\), the problem (\ref{eq:constrained_problem}) has a solution.
-    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
-respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
-    \item If \(k\) from (\ref{eq:points}) is sufficiently large,
-    the values \(\mu_\nu\) from (\ref{step2}) satisfy the inequalities
-    \begin{equation}
-      r_\nu\mu_\nu > 2^{3-1/N}L_\nu \sqrt{N+3},\: 1\leqslant \nu \leqslant m + 1.
-    \end{equation}
-  \end{enumerate}
-  \begin{theorem} (On convergence conditions of MIAGS) Let for each problem \(i,\:1\leqslant i\leqslant q\) from (\ref{eq:many_problems})
-    the conditions 1-3 of the Theorem \ref{th:single_conv} are true, i.e. each problem can be solved by IAGS.
-    Then solving all of the \(q\) problems by MIAGS will
-    generate \(q\) infinite sequences \(\{y^k_i\},\:1\leqslant i\leqslant q\), such that
-    \begin{displaymath}
-      \varphi_i(\overline{y_i})=\inf\{ \varphi(y^k_i): g^i_j(y^k_i)\leqslant 0,1\leqslant j\leqslant m_i, k=1,2,\dots\}=\varphi_i(y^*_i).
-    \end{displaymath}
-  \end{theorem}
 \end{frame}
@@
   \end{columns}
-  \footnotesize{GCGen is available here \textit{https://github.com/UNN-ITMM-Software/GCGen}}
+  \footnotesize{GCGen is available here: \textit{https://github.com/UNN-ITMM-Software/GCGen}}
 \end{frame}
@@
     \begin{itemize}
-      \item implemented support of nonlinear functional constraints in the method
+      \item Implemented support of nonlinear functional constraints in the method
       for solving a set of global optimization problems;
@@
 
-    sovrasov.vlad@gmail.com
+    sovrasov.vladislav@itmm.unn.ru
 

### 2020-10-02 2020/OPTIMA/Answers.tex
OPTIMA: add revised versions of the answers and paper

--- 2020-09-15 2020/OPTIMA/Answers.tex
@@
 
-We thank very much all the reviewers for their useful comments and suggestions.
-Please find below our answers that we suggest to clarify the issues.
+We would like to express our deepest gratitute to all the reviewers for their useful comments and suggestions.
+Please find below our answers and clarifications of the issues.
 
@@
 \textbf{Answer}.
-The paper corrected by an English-speaking person.
+The paper was reviewed and corrected by an experienced translator. Some parts of the text were reformulated or rewritten to improve readability and correct grammatical mistakes.
 
@@
 \textbf{Answer}.
+Clarification of the objective of this work was added at the end of Section 2.
 
@@
 \textbf{Answer}.
-В рассматриваемых задачах целевая функция и ограничения предполагаются многоэкстремальными и заданными в виде ``черного ящика''.
-Кроме того, подразумевается, что каждое испытание (т.е.
-последовательное вычисление значений ограничений и целевой фукнции) является трудоемкой операцией и может потребовать значительных вычислительных ресурсов.
-Такая постановка часто встречается при решении прикладных задач (например, задач нелинейной аппроксимации; решении систем нелинейных уравнений и неравенств; подборе параметров математических моделей по экспериментальным данным, и т.п.).
-Поэтому задачи липшицевой глобальной оптимизации даже с одной переменной до сих пор вызывают интерес исследователей [1,2], а задачи размерности 2-3 являются стандартными для тестирования и сравнения алгоритмов липшицевой глобальной оптимизации [3].
+The target function and constraints in the considered tasks are assumed to be multiextreme and set as a "black box".
+In addition, we assumed that each trial (i.e.
+sequential calculation of constraints and target function values) is a labor-intensive operation and may require significant computational resources.
+Such setting is often encountered in solving applied problems (e.g., problems of nonlinear approximation; solving systems of nonlinear equations and inequalities; selecting parameters of mathematical models from experimental data, etc.).
+Therefore, even the problems of Lipschitz global optimization with only one variable still remain of interest to researchers [1,2], and problems of higher dimensionality (2-3) are considered standard for tests and comparison of algorithms of Lipschitz  global optimization [3].
 
-В статью добавлен комментарий по поводу сложности задач рассматриваемого класса (see ...), а также численные эксперименты
-по решению задач размерности 4 (see Section 4.1 and Table 1).
+We added a comment on the complexity of the considered class of problems (see ...), as well as numerical experiments on solving dimensionality problems 4 (see Section 4.1 and Table 1).
 
@@
 \textbf{Answer}.
-Результаты сравнения core Global Search Algorithm с рядом nature-inspired methods при решении нескольких сотен тестовых задач приведены в [1].
-Результаты показывают, что GSA обеспечивает лучшее качество решения задач (решается большее число задач за меньшее число поисковых испытаний).
-В текст статьи добавлен соответствующий комментарий (see page ...).
+The results of comparison of core Global Search Algorithm with a number of nature-inspired methods when solving several hundred test tasks are given in [1].
+The results show that GSA provides better quality of problem solving (more problems are solved in fewer search trials).
+A corresponding comment has been added to the manuscript (see page 11).
 
-Еще раз отметим, что в работе рассматриваются задачи с функциями вида ``черный ящик'' (i.e., an analytical representation of the functions is not available).
-Поэтому сравнение с методами, использующими аналитическое представление функций задачи (например, основанными на идеях DC-оптимизации, see [2] and references given therein) не будет являться корректным.
-В текст статьи добавлен соответствующий комментарий (see page ...).
+Once again, we would like to point out that the article considers problems with "black box" functions (i.e., an analytical representation of the functions is not available).
+Therefore, comparing them with methods based on analytical representation of the task functions (for example, based on the ideas of DC-optimization, see [2] and references given therein) would not be appropriate.
+An relevant comment has been added to the manuscript (see page 2).
 
@@
 \textbf{Answer}.
-This formula was reformulated.
+This formula was rewritten.
 
@@
 \textbf{Answer}.
-Definitions and terminology in the Section 3 was corrected.
+Definitions and terminology in the Section 3 were corrected.
 
@@
 \textbf{Answer}.
+The presented method converges to all the global optimal solutions at once, but numerically it's hard
+to obtain them all with a guarentee. That happens due to lack of an appropriate stopping criterion.
 
@@
 
+A series of global optimization problems can arise from an mixed-integer problem or
+after scalarization [1] of an multi-objective problem [2]. In both cases, all the problems from the series have to be solved with approximately similar quality, because all of
+them are equally important. To speedup this process under constrained computational resources,
+we can apply parallel computations along with a resource balancing approach to guarantee equal
+quality of solutions in all the problems by the end of the optimization process.
+
+1. Barkalov, K. and I. Lebedev. “Parallel Global Optimization for Non-convex Mixed-Integer Problems.” RuSCDays (2019).
+
+2.
+Jahn J.
+Scalarization in Multi Objective Optimization.
+In: Serafini P.
+(eds) Mathematics of Multi Objective Optimization.
+International Centre for Mechanical Sciences (Courses and Lectures), vol 289.
+Springer, Vienna (1985).
+
 \end{document}

### 2020-10-02 2020/OPTIMA/optima_paper.tex
OPTIMA: fixes, update slides
OPTIMA: add revised versions of the answers and paper

--- 2020-08-26 2020/OPTIMA/optima_paper.tex
@@
 %
-\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
+\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
 \email{sovrasov.vladislav@itmm.unn.ru},
@@
 \begin{abstract}
-In the present work, an extension of global optimization algorithm for solving a set of problems
-simultaneously onto a case of problems with non-convex constraints is considered. The property
-of uniform convergence inherent to the algorithm allows distributing the computation resources
-in the optimal way since the errors of numerical solutions decrease with approximately equal
-rates in all problems in the course of the execution of the algorithm. The algorithm assigns a
-priority to each problem and performs the computations of the objective functions in several
-problems in parallel at every iteration. Upon stop of the method at arbitrary time moment,
-solutions of similar quality would be obtained in all problems of the series being solved. A series
-of several problems arises if the global optimization problem has a discrete parameter or, for
-example, in solving a multicriteria optimization problem by applying scalarization techniques. The
-considered algorithm employs Peano-type space-filling curves for the reduction of the
-multidimensional optimization problems to the one-dimensional ones.  The efficiency of the
-implemented algorithm was tested using the sets of artificially generated global optimization
-problems as well as in solving a series of problems obtained as a result of scalarization of a
-multicriteria optimization  problem. Also, the numerical experiments confirmed the uniform
-convergence of the proposed method.
+This paper proposes an extension of global optimization algorithm for solving a set of problems
+with non-convex constraints. The uniform convergence inherent to the algorithm allows for the optimal
+distribution of the computational resources, since the number of errors in numerical solutions
+decreases at approximately equal rate for all problems processed by the algorithm. The algorithm assigns
+a priority level to each problem and, with every iteration, performs the computations of the objective
+functions for several problems in parallel. After the algorithm is terminated at any given time,
+it obtains solutions of similar quality for all the problems of the set. Such sets appear when
+the global optimization problem has a discrete parameter or, for example, when a multicriteria optimization
+problem is solved by scalarization techniques. The algorithm employs Peano-type space-filling curves
+for the reduction of the multidimensional optimization problems to the one-dimensional ones.
+The efficiency of the implemented algorithm was tested using the sets of artificially generated
+global optimization problems, as well as a series of problems obtained as a result of scalarization
+of a multicriteria optimization problem. Additional numerical experiments also confirmed
+the uniform convergence of the proposed method.
 
@@
 \label{sec-intro}
-Global optimization problems with non-convex constraints are considered to be one
-of the most difficult optimization problems.
-Finding of the global minimum of a function of several variables appears to be more
-difficult often than a local optimization in a thousand-dimensional space.
-For the latter, application of the simplest gradient descent method may appear to be enough,
-whereas to \textit{guarantee} the finding of the global optimum the optimization
-methods have to accumulate the information on the behaviour of the objective function in the
-whole search domain \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}.
-The solving of a series of such problems at limited computational resources is even more
-difficult problem: besides the search of the global extremum, it is necessary to distribute the
-computational resources in such a way that the location of the global extremum would be
-estimated with approximately equal quality in all problems being solved at once.
-Usually, a series of \(q\) problems is solved either sequentially or in parallel by portions of \(p\ll
-q\) problems where \(p\) is a number of parallel computational devices.
-Such an approach leads to a situation when at every time moment until the stop of computations
-some problems remain, in which the estimate of the global optimum is not obtained yet whereas
-in the problems from the beginning of the list the optimum may be estimated even with excess
-precision.
+Global optimization problems with non-convex constraints are considered as one of the most difficult
+optimization problems.
+Finding of the global minimum of a function with several variables often appears to be
+more difficult than a local optimization in a thousand-dimensional space.
+The simplest gradient descent method could be sufficient for the latter, however,
+to \textit{guarantee} that the global optimum is found, the optimization
+methods have to accumulate the information on the behavior of the objective function
+in the entire search domain  \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}.
+In some cases, for instance, for DC or QP problems, more effective methods and even
+analytical optimality conditions could be applied  \cite{PhamDinh2014, qpBook},
+but these methods are barely applicable for black box functions (i.e., an analytical
+representation of the functions is not available).
 
-In the present work, a generalization of parallel global optimization method for simultaneous
-solving of a set of problems developed earlier in Lobachevsky State University of Nizhni Novgorod
-\cite{BarkalovStrongin2018} onto the case of non-convex constraints is considered. To account
-for the constraints, the index scheme \cite{Strongin2000,Pugliese} is applied allowing working with
-partially defined objective function and constraints and featured by efficiency comparable to other
-approaches \cite{BarkalovLebedev2017}. The convergence of the algorithm to the global
-optimizers in all problems was proved theoretically, the efficiency of the implemented
-algorithm was demonstrated by solving sets of problems generated by a special mechanism
-generating the sets of problems of predefined dimensionality with predefined number of
-non-convex constraints \cite{GergelBarkalov2019}.
-Besides the problems generated artificially, the considered method was tested also on a set of
-problems arising when solving a multicriteria optimization problem with nonlinear constraints by
+Solving a series of such problems with limited computational resources is even more
+difficult: besides the search of the global extremum, the computational resources
+have to be distributed in such a way that the location of the global extremum would be
+estimated with approximately equal quality for all problems being solved at the same time.
+Usually, a series of  \(q\) problems is solved either sequentially or in parallel by sets of  \(p\ll
+q\) problems where \(p\) is the number of parallel computational devices.
+Such an approach creates a situation when at every given moment there exist problems for which
+the global optimum is not obtained yet, while the optimum for the problems from the beginning of
+the set may be estimated even with excess precision.
+
+This paper considers a parallel global optimization method developed at the Lobachevsky
+State University of Nizhny Novgorod for simultaneous solving of a set of problems \cite{BarkalovStrongin2018}
+and in particular, its generalization to cover the non-convex constraints. The index scheme \cite{Strongin2000,Pugliese}
+which is applied to account for the constraints, allows operating with partially defined objective functions and constraints, and its efficiency is on par with similar approaches \cite{BarkalovLebedev2017}.
+The convergence of the algorithm to the global optimizers in all problems was proved theoretically,
+the efficiency of the implemented algorithm was demonstrated on sets of problems generated
+by a special mechanism. This mechanism generates sets of problems with predefined dimensionality
+and predefined number of non-convex constraints  \cite{GergelBarkalov2019}.
+Besides the artificially generated problems, the method was also tested on sets of
+multicriteria optimization prob¬lems with nonlinear constraints solved by
 criteria convolution method  \cite{Ehrgott2005}.
 
-The paper is structured as follows. In Section \ref{sec:problem}, the statement of the problem to
-be solved is considered. Then, in Section \ref{sec:method}, a description of the parallel
-optimization method is given. The sufficient conditions of convergence of the considered
-method are given in Section \ref{sec:conv_method}.
+The paper is structured as follows. Section \ref{sec:problem} provides the statement of the problem.
+Section \ref{sec:method} gives a description of the parallel optimization method.
+Subsection \ref{sec:conv_method} describes the sufficient conditions of convergence
+of the considered method.
 The results of the numerical experiments confirming the efficiency of considered method are
 presented in Section \ref{sec:exps}.
-In Conclusion, a short summary of the results obtained in the present study is given and the
-directions of further effort on the improvement of the software implementation of the
-considered method are outlined.
+Conclusion gives a short summary of the results of the study and proposes the
+directions for further improvement of the software implementation of the considered method.
 
-\section{Global optimization problem statement}
+\section{Statement of the global optimization problem}
 \label{sec:problem}
-In this work we consider the following
-problem: to find the global minimum of an \(N\)-dimensional function \(\varphi(y)\) in
+This paper considers the following problem:
+to find the global minimum of an  \(N\)-dimensional function \(\varphi(y)\) in
 a hyperinterval $D$
@@
 \end{displaymath}
-In order to construct an estimate of the global minimum based on a finite number of
-computations of the function values, the rate of variation of \(\varphi(y)\) in \(D\) is required to
+In order to produce an estimate of the global minimum based on a finite number of
+computations of the function values, the rate of variation of \(\varphi(y)\) in \(D\) has to
 be limited.
@@
 
-There are various methods solving the considered multidimensional problem directly
+There are various methods for solving the considered multidimensional problem directly
 \cite{SergeyevKvasov2017, Jones2009} as well as efficient methods for solving the univariate problems \cite{Norkin1992, Strongin2000}.
-In the present work, a one-dimensional method is considered, which is applied together with the
-dimensionality reduction scheme.
-The use of space-filling curve (or \textit{evolvent}) $y(x)$, where
+This paper considers a one-dimensional method, which is applied jointly with
+the dimensionality reduction scheme.
+The use of space¬filling curve (or \textit{evolvent}) $y(x)$, where
 \begin{equation}
@@
 \end{equation}
- is the well-known scheme of dimensionality reduction
+ is a well-known scheme of dimensionality reduction
  of the initial problem for the global optimization algorithms
@@
 A mapping of the type (\ref{cube}) allows reducing a multidimensional problem to a univariate one at the expense of worsening its properties. In particular, the
-one-dimensional function \(\varphi(y(x))\) is not lipschitzian but h\"{o}lderian one:
+one-dimensional function \(\varphi(y(x))\) is not a Lipschitz function but a Holder one:
 \begin{displaymath}
@@
 \end{displaymath}
-where the H\"{o}lder constant \(H\) is related to the Lipschitz one \(L\) by the relation
+where the H\"{o}lder constant \(H\) is related to the Lipschitz constant \(L\) as
 \begin{displaymath}
@@
 
-Let us denote \(g_{m+1}(y)=\varphi(y)\). Hereafter, we will assume all functions
+Let us set \(g_{m+1}(y)=\varphi(y)\). Hereafter, we shall assume all functions
 \(g_k(y),1\leqslant k \leqslant m+1\),
-to satisfy the Lipschitz condition in some hyperinterval including \(D\).
+to satisfy the Lipschitz condition in a certain hyperinterval including \(D\).
 
@@
 
+Similar to \cite{BarkalovStrongin2018}, this work aims to formulate a method that will ensure a uniform convergence of the
+solutions of all problems in the series. Here and below, a uniform convergence implies a proportional decrease
+of the distance between the best iteration point and the
+global optimizer in all the problems in the series:
+\begin{equation}
+  \exists \varepsilon > 0: \forall s>1, \forall i,j\in\{1,\dots,q\}\;
+    \frac{\Vert \tilde{y_i}(s)^* - y_i*\Vert_\infty}{\Vert \tilde{y_j}(s)^* - y^*_j\Vert_	\infty} \leqslant \varepsilon,
+\end{equation}
+where \(s\) is the number of steps of the optimization method, \(\tilde{y_i}(s)^*\) is the best iteration point
+in the problem \(i\) from the set (\ref{eq:many_problems}) at the step \(s\).
+
 \section{Description of the global optimization method}
@@
 
-Taking into account the dimensionality reduction scheme (\ref{cube}), let us assume when
-describing the method that it is required to find the global minimum of the function \(\varphi(x),
-x\in[0,1]\), satisfying the H\"{o}lder condition with the constraints \(g_j(x)\) also satisfying this
-condition in the interval \([0,1]\).
+Taking into account the dimensionality reduction scheme (\ref{cube}), let us assume that the method
+requires finding the global minimum of the function \(\varphi(x),
+x\in[0,1]\), which satisfies the H\"{o}lder condition with the constraints \(g_j(x)\), which in turn satisfy
+this condition in the interval \([0,1]\).
 
-The index algorithm of global search (IAGS) for solving the one-dimensional problems
+The index algorithm of global search (IAGS) for solving one-dimensional problems
 %(\ref{eq:constrained_problem})
-considered here implies the constructing of a sequence of points
-\(x_k\), which the values of index function \(z_k\) are computed in.
-On account of, the index scheme \cite{Strongin2000} was used.
-Let's assume that each function \(g_i(x), 1 \leqslant i \leqslant m + 1\), is defined and computable
+considered here implies construction of a sequence of points
+\(x_k\), at which the values of index function \(z_k\) are calculated.
+The index scheme \cite{Strongin2000} was used to account for the latter.
+Let us assume that each function \(g_i(x), 1 \leqslant i \leqslant m + 1\), is defined and computable
 only in the corresponding subrange \(Q_i \in [a, b]\), where \(\varphi(x)\) is denoted as \(g_{m+1}(x)\) and
@@
 \end{equation}
-In the view of the definition (\ref{eq:q}) the initial problem can be rewritten in the form
+Considering the definition (\ref{eq:q}) the initial problem can be rewritten as
 \begin{displaymath}
@@
 
-The index scheme assigns a number called index \(\nu(x)\) to the each point of the search sequence \(x_k\):
+The index scheme assigns a number called index \(\nu(x)\) to each point of the search sequence \(x_k\):
 \begin{equation}
@@
 
-In order to obtain an index of the point \(x\) we have to perform a trial defined via the following steps:
+In order to obtain the index of the point \(x\) we have to perform a trial defined via the following steps:
 \begin{enumerate}
-  \item Consequently compute functions \(g_i(x),1 \leqslant i \leqslant m\) until for some \(i^*\) \(g_{i^*}(x)>0\) or until \(i=m+1\).
+  \item Sequentially  compute functions \(g_i(x),1 \leqslant i \leqslant m\) until for some \(i^*\) \(g_{i^*}(x)>0\) or until \(i=m+1\).
   Set \(\nu(x)=i^*\) or \(\nu(x)=m+1\) if \(g_i(x) \leqslant 0, 1 \leqslant i \leqslant m\).
   Therefore, \(\nu(x)\) is the number of the first violated constraint at the point \(x\) or \(m+1\) if
-  all constraints are satisfied at \(x\).
-  \item Return pair \(\nu(x), z=g_{\nu(x)}(x)\) as a result of trial.
+  all the constraints are satisfied at \(x\).
+  \item Return pair \(\nu(x), z=g_{\nu(x)}(x)\) as a result of the trial.
 \end{enumerate}
 
-Such an approach to conducting the trials allows reducing the initial problem with functional
+This approach to trials allows reducing the initial problem with functional
 constraints to an unconstrained problem of minimization of a discontinuous function:
@@
 Despite the values of the H\"{o}lder constants \(H_k\) and the value \(g_{M}^{*}\) are not
-known in advance, these ones can be estimated in the course of solving the problem.
+known in advance, they can be estimated in the course of solving the problem.
 A set of triples \(\{(x_k,\nu_k,z_k)\}, 1\leqslant k\leqslant n\), constitutes the search information
-accumulated by the method upon executing \(n\) steps.
+accumulated by the method upon execution of \(n\) steps.
 
@@
 the interval \([0,1]\).
-The indices of the points 0 and 1 are considered to be zero ones, the values \(z\) in these points
+The indices of the points 0 and 1 are considered to be zero indices, the values \(z\) at these points
 are undefined.
-Let \(k\geqslant 1\) iterations of the method have been performed, in the course of which the
-trials were conducted in \(k\) points \(x_i, 1\leqslant i\leqslant k\).
+Let us assume that \(k\geqslant 1\) iterations of the method have been performed. In the course of this performance the
+trials were conducted at \(k\) points \(x_i, 1\leqslant i\leqslant k\).
 Then, the points \(x^{k+1}\) of the search trials of the next \((k+1)^{\rm th}\) iteration are
@@
 
-Step 1. Renumber the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which
-includes the boundary points of the interval \([0,1]\) as well as the points of preceding trials by
-the lower indices in the order of increasing coordinate values, i.e.
+Step 1. Reassign lower indices to the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which
+includes the boundary points of the interval \([0,1]\) as well as the points of preceding trials in the order of increasing coordinate values, i.e.
 \begin{equation}
@@
 \end{equation}
-and juxtapose to them the values \(z_{i}=g_{\nu }(x_{i}),\nu =\nu (x_{i}),i={\overline {1,k}}\) computed at these points.
+and compare them with the values \(z_{i}=g_{\nu }(x_{i}),\nu =\nu (x_{i}),i={\overline {1,k}}\) computed at these points.
 
 Step 2. For each integer number \(\nu ,1\leqslant \nu \leqslant m+1\), determine the corresponding
-set \(I_{\nu }\) of the lower indices of the points, which the values of the functions \(g_{\nu
-}(x)\) are computed in:
+set \(I_{\nu }\) of the lower indices of the points, at which the values of the functions \(g_{\nu
+}(x)\) are computed:
 \begin{displaymath}
@@
 \end{equation}
-If the set \(I_{\nu }\) contains less than two elements or if the value \(\mu _{\nu }\) appears to
-be equal to zero, then accept \(\mu _{\nu }=1\).
+If the set \(I_{\nu }\) contains less than two elements or if the value \(\mu _{\nu }\) is equal to zero, then assume \(\mu _{\nu }=1\).
 
-Step 4. For all nonempty sets \(I_{\nu },\nu ={\overline {1,M}}\) compute the estimates
+Step 4. For all the nonempty sets \(I_{\nu },\nu ={\overline {1,M}}\) compute the estimates
 \begin{equation}
@@
 The values \(r_{\nu }>1,\nu ={\overline {1,m}}\) are the parameters of the algorithm.
-The products \(r_{\nu}\mu_{\nu}\) used in computing the characteristics as the estimates of the
-unknown H\"{o}lder constants depend on these ones.
+They define the products \(r_{\nu}\mu_{\nu}\) used in computing the characteristics as the estimates of the
+unknown H\"{o}lder constants.
 
@@
 
-The algorithm stops if the condition \(\Delta_{t}\leqslant \varepsilon\) is satisfied where
+The algorithm stops if the condition \(\Delta_{t}\leqslant \varepsilon\) is satisfied. Here
 \(\varepsilon>0\) is a predefined precision.
@@
 \end{equation}
-are accepted as the estimates of the global solution.
+are assumed as the estimates of the global solution.
 
-Next, following the approach described in \cite{BarkalovStrongin2018}, we will use \(q\) copies
+Next, following the approach described in \cite{BarkalovStrongin2018}, we shall use \(q\) copies
 of IAGS working synchronously to solve the problem series (\ref{eq:many_problems}).
 The only difference is that when selecting the interval with the maximum characteristic at Step
-6, the choice will be made from all intervals, which \(q\) copies of IAGS have generated to the
-moment.
+6, the choice will be made from all intervals, which were generated by \(q\) copies of IAGS.
 If the maximum characteristic corresponds to the \(i^{\rm th}\) problem, then Step 7 is
-performed in the \(i^{\rm th}\) copy of the method while the rest copies stay idle.
-This way, at every iteration the trial is performed in the most promising problem from the
-viewpoint of the characteristics (\ref{step3_1}) that allows distributing the resources of the
+performed in the \(i^{\rm th}\) copy of the method while the other copies stay idle.
+This way, at every iteration the trial is performed for the most promising problem in terms of the characteristics (\ref{step3_1}). This allows distributing the resources of the
 method among the problems dynamically. Let's denote this method as MIAGS.
 
-The parallel modification of the method doesn't differ from the one considered in
-\cite{BarkalovStrongin2018} and consists in the selection of \(p\) intervals at Step 6 and the
-performing of \(p\) trials in parallel at the next step.
-At that, all resources of the method within the framework of the iteration may be directed onto
-a single problem as well as onto \(l\leqslant p\) problems simultaneously (subject to the problem,
-which the intervals selected by the method belong to).
+The parallel modification of the method does not differ from the one considered in
+\cite{BarkalovStrongin2018} and consists in the selection of \(p\) intervals at Step 6 and
+performing  \(p\) trials in parallel at the following step.
+All resources of the method within the framework of the iteration may be focused on
+a single problem as well as on \(l\leqslant p\) problems simultaneously (depending on the problem,
+to which the intervals selected by the method belong).
 
@@
 
-The conditions of convergence of the described in Section \ref{sec:method} method in case of \(q=1\) are given in \cite{Strongin2000}.
+The conditions of convergence of the method described in Section \ref{sec:method} in case of \(q=1\) are given in \cite{Strongin2000}.
 \begin{theorem} (Sufficient convergence conditions)
   \label{th:single_conv}
-  Assume that the following conditions are true:
+  Let us assume that the following conditions are true:
   \begin{enumerate}
     \item \(D\ne\emptyset\), the problem (\ref{eq:constrained_problem}) has a solution.
-    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
+    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitz functions with
 respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
@@
   Then any limit point \(\overline{y}\) of the sequence \(\{y_k\} = \{y(x_k)\}\) generated
-  by the index method while solving the problerm
+  by the index method, the problem
   (\ref{eq:constrained_problem}) is feasible and satisfies the conditions
@@
 
-\begin{theorem} (On convergence conditions of MIAGS) Let for each problem \(i,\:1\leqslant i\leqslant q\) from (\ref{eq:many_problems})
-  the conditions 1-3 of the Theorem \ref{th:single_conv} are true, i.e. each problem can be solved by IAGS.
+\begin{theorem} (On the convergence conditions of MIAGS) Let the conditions 1-3 of the Theorem \ref{th:single_conv} for each problem \(i,\:1\leqslant i\leqslant q\) from (\ref{eq:many_problems}), be true
+i.e. each problem can be solved by IAGS.
   Then solving all of the \(q\) problems by MIAGS will
@@
   where \(k\) corresponds to the number of trials in the first problem and \(s\)
-  corresponds to the number of trials in the second one. The sequence of trials \(\{v^k\}\)
+  corresponds to the number of trials in the second problem. The sequence of trials \(\{v^k\}\)
   corresponds to the first problem and the sequence of trials \(\{u^s\}\)
   corresponds to the second problem. The values \(z^k=g^\varphi_\nu(v_k),\nu =\nu (v_{k})\) correspond to the
-  trial points \(\{v^k\}\) and the values \(w^s=g^\psi_\nu(u_s),\nu =\nu(u_{s})\) correspond to the
+  trial points \(\{v^k\}\), and the values \(w^s=g^\psi_\nu(u_s),\nu =\nu(u_{s})\) correspond to the
   trial points \(\{u^s\}\).
 
-  When the considered two problems are solved simultaneously, the algorithm selects
+  When the two problems are solved simultaneously, the algorithm selects
   an interval for the next trial according to the condition:
@@
 
-  Let the algorithm is solving two problems and the iterations counter is \(l = k + s, l=0,1,2,\dots\).
-  Then, since for all of the problems Theorem \ref{th:single_conv} is true, at least one of the sequences
-  \(\{v^k\}\) and \(\{u^s\}\) will be infinite (let it be \(\{v^k\}\)). If we prove that both the
-  sequences are infinite, this would lead us to convergence in both the considered problems.
+  Let the algorithm solve two problems and the iterations counter be \(l = k + s, l=0,1,2,\dots\).
+  Then, since Theorem \ref{th:single_conv} is true for all the problems , at least one of the sequences
+  \(\{v^k\}\) and \(\{u^s\}\) will be infinite (let it be \(\{v^k\}\)). If we can prove that both   sequences are infinite, this will demonstrate the convergence in both considered problems.
 
-  Let the limit point \(\overline{v}\in [v_{i-1},v_i]\), where \(i=i(k)\). The indexes of \(v_{i-1}\;,v_i\)
-  can be equal or different, but, because of convergence, starting from large enough \(k\) they will be stable.
+  Let us take a limit point \(\overline{v}\in [v_{i-1},v_i]\), where \(i=i(k)\). The indices \(v_{i-1}\;,v_i\)
+  can be equal or different, but, because of convergence, if \(k\) is large enough they will be stable.
   %will be steady  equal to \(m_1+1\) (if \(\overline{v}\) is
@@
   \end{displaymath}
-  Considering that, we can establish an upper bound:
+  Having considered that we can establish an upper bound:
   \begin{displaymath}
@@
   \end{displaymath}
-  Therefore using the first branch form rule (\ref{step3_1}) we have inequality
+  Therefore using the first branch of form rule (\ref{step3_1}) we get an inequality
   \begin{equation}
@@
   \end{equation}
-  Otherwise because of \(\overline{v}\) is a feasible point \(z_i\to 0\) at \(k\to\infty\).
+  Otherwise because  \(\overline{v}\) is a feasible point \(z_i\to 0\) at \(k\to\infty\).
 
-  From (\ref{eq:th1}) (\ref{eq:th2}) (\ref{eq:th3}) for any small \(\delta > 0\) exists a large value of \(k\) such that
+  From (\ref{eq:th1}) (\ref{eq:th2}) (\ref{eq:th3}) for any small \(\delta > 0\) there exists a large value of \(k\) such that
   \begin{equation}
@@
 
-  Let \(\alpha = \max\{\nu(u):u\in\{u^s\}\}\). Because \(\alpha\) is currently the greatest index
+  Let \(\alpha = \max\{\nu(u):u\in\{u^s\}\}\). Because \(\alpha\) is currently the highest index
   in the search sequence \(\{u^s\}\) and according to the rule (\(\ref{eq:step_4}\)), \(\exists j: w^*_\alpha=w_j\).
@@
   \end{equation}
-  Taking into account the Remark \ref{rem:r1}, (\ref{eq:th4}), (\ref{eq:th5}) we can conclude that
-  exists such interval that \(R_\psi(j)>0\). At the same time, (\ref{eq:th6}) is true and
+  Taking into account the Remark \ref{rem:r1}, (\ref{eq:th4}), (\ref{eq:th5}) we can conclude that such an interval
+  exists that \(R_\psi(j)>0\). At the same time, (\ref{eq:th6}) is true and
   the inequality
@@
   \end{displaymath}
-  will be true at large enough value of \(k\). Thus the next scheduled iteration is performed on
+  will be true when the value \(k\) is large enough. Thus the next scheduled iteration is performed for
   the second problem with objective \(\psi(y)\), i.e. sequence \(\{v^s\}\) will be infinite as well.
@@
   Since we considered two arbitrary problems from the given set of \(q\) problems, the theorem is
-  true for any pair of problems from the set. By induction the theorem is also true for the whole
+  true for any pair of problems from the set. By induction, the theorem is also true for the whole
   set.
@@
 
-The use of set of test problems with known solutions generated by some random mechanisms is
-one of commonly accepted approaches to the comparing of the optimization algorithms
+The use of set of test problems with known solutions generated by a random mechanisms is
+one of the commonly accepted approaches to the comparison of optimization algorithms
 \cite{Beiranvand2017}.
-In the present work, two generators of test problems generating the problems of different nature
+Experiments presented in this paper were based on two generators of problems of different nature
 \cite{grishaginClass, Gaviano2003} were used.
-These generators generate the problems without the nonlinear constraints. Therefore, the
+These generators produce problems without nonlinear constraints. Therefore, the
 GCGen\footnote{The original code of the system can be accessed at https://github.com/UNN-
-ITMM-Software/GCGen} \cite{GergelBarkalov2019} system was used in addition.
-This system allows generating the problems with constraints based on arbitrary nonlinear
+ITMM-Software/GCGen} \cite{GergelBarkalov2019} system was to be used to supplement these generators.
+This system allows generating problems with constraints based on arbitrary nonlinear
 functions.
 
-Along with GCGen system, the examples of its application and constructing the set of problems
-each consisting of an objective function and two constraints generated by \(F_{GR}\)
-\cite{grishaginClass} generator or GKLS \cite{Gaviano2003} one Are distributed.
+The GCGen system comes with the examples of its application and construction of sets of
+problems each consisting of an objective function and two constraints generated by \(F_{GR}\)
+\cite{grishaginClass} or GKLS \cite{Gaviano2003} generator.
 GKLS \cite{Gaviano2003} allows obtaining the functions of predefined dimensionality and
@@
   two-dimensional GKLS functions and 50 problems with \(F_{GR}\).
-  This set is generated in order to demonstrate the properties of the method to be preserved at essentially
-  different properties of the problems.
+  This set is generated in order to demonstrate that the efficiency of the method is preserved at essentially
+  varying properties of the problems.
 \end{itemize}
@@
 Examples of contour plots of the considered problems are presented in Fig. \ref{fig:isolines}.
-The feasible domain is highlighted by color.
+The feasible domain is highlighted.
 
@@
 where \(a\) and \(b\) are the left and right boundaries of the hypercube from (\ref{eq:task}).
-If this relation is not fulfilled upon reaching the limit of number of iterations, the problem was
+If this relation is not fulfilled after the limit of number of iterations had been reached, the problem was
 considered to be not solved.
 
-When evaluating the quality of the method and its implementation, besides the speedup due to
-the parallelization and the work time,
-Was taken into account as well.
-The mean maximum distance (in terms of \(l_{\infty}\)-norm) form the current estimate of the
+When evaluating the quality of the method and its implementation, besides increased
+computation speed due to parallelization,we shall also account for the mean maximum
+distance (in terms of \(l_{\infty}\)-norm) from the current estimate of the
 optimum to its actual position computed on the set of problems (\ref{eq:many_problems}):
@@
 
-The parallel method was implemented on C++ language with the use OpenMP technology for
-the parallelization of the trial execution process in the shared memory.
+The parallel method was implemented in  C++  with the use OpenMP technology for
+parallelization of the trial execution process in the shared memory.
 All numerical experiments were carried out using a computer with the following configuration:
@@
 
-\subsection{Result of solving the generated problems}
+\subsection{Solution results of generated problems}
 
-The results of solving the test problems by the sequential and parallel versions of the
+The solution results of test problems by the sequential and parallel versions of the
 MIAGS are presented in Table~\ref{tab:speedup}.
 For all two-dimensional classes of problems, the parameter \(r=4.7\).
-In the case of the three and four-dimensional problems, \(r=4.7\) and also to improve the convergence speed
+In the case of the three and four-dimensional problems, \(r=4.7\).  The convergence speed
+was improved by applying
 \(\varepsilon\)-reservation technique from \cite{Strongin2000} Chapter 8.3 with
-\(\varepsilon=0.1\) has been employed.
-In all experiments, an additional computational load was introduced into the objective function
-and constraints so that the time of a single call of a problem function was equal to
+\(\varepsilon=0.1\).
+In all the experiments, an additional computational load was introduced into the objective function
+and constraints to get the duration of a single call of a problem function equal to
 approximately 1 ms.
 
-One can see from Table \ref{tab:speedup} that the speedup in the iterations \(S_i=\frac{iters(p=1)}{iters(p=i)}\) increased
-linearly with increasing number of threads \(p\) whereas the speedup in time \(S_t=\frac{time(p=1)}{time(p=i)}\) increased
-not so fast that points to a non-ideal implementation of the algorithm.
-One can increase the actual speedup, the upper limit for which is \(S_i\) by the optimization of
-the interaction between the copies of IAGS. It is planned to do in the course of future work.
+Table \ref{tab:speedup} shows that the speedup in the iterations \(S_i=\frac{iters(p=1)}{iters(p=i)}\) increased
+linearly with increasing number of threads \(p\), whereas the speed of calculations \(S_t=\frac{time(p=1)}{time(p=i)}\) increased
+at a lower rate, which points to a non-ideal implementation of the algorithm.
+The actual speedup, the upper limit for which is \(S_i\), can be increased by the optimization
+of the interaction between the copies of IAGS. We plan to test this approach in our future works.
 
@@
 In order to demonstrate the uniform convergence of MIAGS, all test problems have been solved by IAGS as well.
+IAGS is comparable with other stochastic and deterministic derivative-free global optimization algorithms \cite{Sovrasov2019}. Thus, it
+provides a strong baseline in solving a single global optimization problem.
 Fig. \ref{fig:devs_mixed} presents the plots of mean and maximum distances from the actual
-optima to the current estimates of these ones when solving a series of problems generated by
-two different generators separately (solid line) and together (dashed line).
-In spite of an essential differences in the structure of problems, the MIAGS decreased
-the maximum deviations of the estimates from the optima as well as the mean ones much faster.
-It evidences the uniform convergence over the whole set of problems being solved all together.
-At that, in the case of the sequential solving of the problems, the magnitude \(D_{max}\) takes
+optima to the current estimates of the optima when solving a series of problems generated by
+two different generators separately (solid line) and jontly (dashed line).
+Despite the essential differences in the structure of problems, the MIAGS decreased
+the maximum deviations of the estimates from the optima, as well as the mean optima much faster.
+It evidences that the uniform convergence over the whole set of problems has been achieved.
+In the case of the sequential solving of the problems, the magnitude \(D_{max}\) takes
 its maximum value until the last problem is solved.
@@
 
-\subsection{Example of solving a multicriteria problems}
+\subsection{Example of solving a multicriteria problem}
 
-In order to demonstrate the efficiency of the approach to the balancing of the load, let us
+In order to demonstrate the efficiency of the approach in the balancing of the load, let us
 consider an example, in which a set of problems of the kind (\ref{eq:many_problems}) is
@@
 where \(\lambda_1,\lambda_2\in[0,1],\: \lambda_1+\lambda_2=1\).
-Trying all possible convolution coefficients, one can find the whole set of Pareto-optimal
-solutions in the problem (\ref{eq:mco_probem}).
+Testing all possible convolution coefficients allows finding the whole set of Pareto-optimal
+solutions of the problem (\ref{eq:mco_probem}).
 For the numerical construction of the Pareto set, let us select 100 sets of coefficients
@@
 
-As the limitation for the computational resources, a limit of 2500 trials was selected.
+Computational resources were limited to 2500 trials.
 The set of auxiliary scalar problems was solved by two methods:
@@
 the computational resources were distributed among the problems uniformly;
-  \item all problems are solved simultaneously using MIAGS with a preset limit of
+  \item all problems were solved simultaneously using MIAGS with a preset limit of
 2500 trials.
@@
 All plots agree with the ones presented in \cite{BinhKorn1999} qualitatively (the authors
-haven't provided any other information to compare).
-One can see the Pareto curve in Fig. \ref{fig:mco_pareto_1} to have concavities that doesn't
-match the solution presented in \cite{BinhKorn1999} and means the lack of resources for
-solving some auxiliary problems.
+did not provide any other information to compare).
+The Pareto curve in Fig. \ref{fig:mco_pareto_1} has concavities that do not
+match the solution presented in \cite{BinhKorn1999}, which means there are not enough resources for solving some auxiliary problems.
 To evaluate the quality of solution, the index \(Spacing(SP)\) \cite{RiquelmeLucken2015}
-featuring the density of the points approximating the Pareto set was computed also.
+featuring the density of the points approximating the Pareto set was computed.
 \begin{equation*}
@@
 \end{equation*}
-In the case of separate solving of the problem \(SP_{single}=0.984\) whereas when solving of
-the problems by the method with the balancing of the load \(SP_{multi}=0.749\) that evidences
-a better quality of the approximation of the solution.
+When problems were solved separately  \(SP_{single}=0.984\), when the load balancing method was
+applied  \(SP_{multi}=0.749\), which evidences a better quality of the approximation
+of the solution.
 
@@
 
-In the course of present work, the support of the non-convex constraints was implemented in the
-algorithm solving a set of the global optimization problems.
-Also, the sufficient conditions of convergence for the developed method were obtained.
-The numerical experiments demonstrating the advantages of the considered approach over
-solving the problems separately were carried out.
-The efficiency of solving a set of problems together was demonstrated by the example of
-solving a multicriteria problem with nonlinear constraints.
-
-In the course of further work, it is planned to improve current implementation of the algorithm
-by reducing the costs of supporting the search information for the set of problems and, this way,
-improving the indices of time speedup due to the parallelization.
-Also, it is planned to implement a version of considered algorithm working in the distributed
-memory according to the scheme described in \cite{BarkalovLebedev2017_2}.
+This paper demonstrates how the support of the non-convex constraints can be implemented
+in the algorithm to solve a set of the global optimization problems.
+This study allowed to find the sufficient conditions of convergence for the developed method.
+The numerical experiments conducted within this research demonstrate the advantages
+of the considered approach over separate solution of the problems.
+The efficiency of joint solution of a set of problems was demonstrated on the example a
+multicriteria problem with nonlinear constraints.
+Further research in this direction should improve current implementation
+of the algorithm by reducing the support costs of the search information for
+the set of problems. This, in turn, should improve the calculation speed
+due to parallelization. There are also plans to implement a version of considered algorithm in the distributed memory according to the scheme described in  \cite{BarkalovLebedev2017_2}.
 

### 2020-10-04 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
Reference to the new grant

@@
 
-\title{Computationally Efficient Approach for Solving Lexicographic Multicriteria Optimization Problems\thanks{This research was supported by the Russian Science Foundation, project No 16-11-10150.
+\title{Computationally Efficient Approach for Solving Lexicographic Multicriteria Optimization Problems\thanks{The work was supported by the Ministry of Science and Higher Education of the Russian Federation, agreement No 13.1902.21.0026.
 %``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''.

### 2020-10-05 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
intermidiate version of answers

@@
 
-%\usepackage{hyperref}
+\usepackage{hyperref}
+\usepackage{cite}
+
 %\usepackage{academicons}
@@
 The further structure of the paper is as follows.
-Section \ref{sec:2} presents a new class of optimization problems–multistage lexicographic multicriteria optimization (MMCOlex) problems, the solution of which is reduced to solving a sequence of global optimization problems with nonlinear constraints.
+Section \ref{sec:2} presents a new class of optimization problems -- multistage lexicographic multicriteria optimization (MMCOlex) problems, the solution of which is reduced to solving a sequence of global optimization problems with nonlinear constraints.
 Section \ref{sec:3} considers the basics of the developed approach: reducing multidimensional MCOlex problems to one-dimensional global optimization problems, applying efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization, and reusing the search information obtained in the process of calculations.
@@
 \begin{equation}\label{eq:10}
-\mathbb{P}_t=\{ P_1,P_2,\dots,P_t\},
+%\mathbb{P}_t=\{ P_1,P_2,\dots,P_t\},
+\mathbb{F}=\{ F_1,F_2,\dots,F_t\},
 \end{equation}
-which can be changed dynamically during calculations by adding new or deleting existing multicriteria optimization problems.
-In general, this approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex).
+%new text begin
+where $F_t, t>1,$ is a MCOlex problem. The size of the set $\mathbb{F}$ can be changed 
+dynamically by adding new required MCOlex problems or removing unnecessary ones. 
+The problems of the set $\mathbb{F}$ specify various requirements for the desired solutions and, thus, all problems of the set $\mathbb{F}$ should be solved in the course of calculations.
+In general, this approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex)\footnote{From a general point of view, the MCOlex problem can be considered as a special case of the MMCOlex problem for $t=1$.}.
+%new text end
 
@@
 However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}.
+%new text begin
+This possibility of reusing search information can be briefly presented as follows.
+%\begin{itemize}
+%\item 
+%При минимизации очередного критерия $f_l(y(x)), 1\leq l\leq s,$ получаемая поисковая информация накапливается в форме множества $A_k$ из (\ref{eq:21}) (следует отметить, что в $A_k$ наряду со значениями $z_i, 1\leq i\leq k,$ минимизируемого критерия $f_l(x)$ запоминаются и вычисленные значения критериев $f_i, 1\leq i\leq k$.
+%\item 
+%При переходе к минимизации следующего критерия $f_{l+1}(y(x))$ значения этого критерия $z_i, 1\leq i\leq k,$ во всех точках $x_i, 1\leq i\leq k,$ ранее выполненных итераций глобального поиска могут получены при помощи быстро выполнимых вычислений на основе запомненных значений критериев $f_i, 1\leq i\leq k$ в поисковой информации $A_k$
+%\begin{equation}\label{new_equation}
+%z'_i=Z(f_i), 1\leq i\leq k,
+%\end{equation}
+%где процедура вычисления значений $z'_i, 1\leq i\leq k$ определяется правилами (\ref{eq:9}), (\ref{eq:18})--(\ref{eq:20}) без повторных вычислениях значений критериев $f_l(y(x)), 1\leq l\leq s$.
+%\end{itemize}
+%new text end
 This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of numerical experiments.
 
-Moreover, using the search information $A_k$ allows reducing the multistage solution of the MCO problem in accordance with the scheme (\ref{eq:9}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:9}), namely
+Moreover, using the search information $A_k$ allows reducing the multistage solution of the MCOlex problem in accordance with the scheme (\ref{eq:9}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:9}), namely
 \begin{equation}\label{eq:30}
@@
 \end{equation}
-
+%new text begin
+%При этом, как и ранее, при любых изменениях оценок  $f_j^*, \delta_j, g_j^k, 1\leq j < s, 1 \leq i \leq k,$ решение задачи (\ref{eq:30}) может быть продолжено с использованием всей имеющейся поисковой информации после быстро выполняемых вычислений новых значений $z'_i,1\leq i \leq k,$ в соответствии с процедурой (\ref{new_equation}).
+%new text end
 In particular, a similar one-step scheme (\ref{eq:30})--(\ref{eq:31}) can be applied through the use of AGCS.
@@
 when at the iteration points $x_i$, $1 \leq i \leq k$, the values of all criteria $f_j(x_i)$, $1 \leq j \leq s$, $1 \leq i \leq k$, are calculated, the effect of using the search information $A_k$ can be even more significant.
-In this case, the search information can also be used when changing the values of the concessions $\delta_j$, $1 \leq j < s$, and thus, the solution of the next problem from the $\mathbb{P}_t$ family (\ref{eq:10}) can be carried out repeatedly using the results of all previously performed calculations.
+In this case, the search information can also be used when changing the values of the concessions $\delta_j$, $1 \leq j < s$, and thus, the solution of the next problem from the $\mathbb{F}$ family (\ref{eq:10}) can be carried out repeatedly using the results of all previously performed calculations.
 The results of numerical experiments (see Section \ref{sec:5}), demonstrate that the volume of computations performed, based on reusing search information, can significantly decrease.
 
-The AGCS algorithm, supplemented by the ability to use search information to solve multiple MCOlex problems, will be referred to below as the global search algorithm for multistage solutions of a set of multiextremal optimization problems with nonlinear constraints (MAGCS).
+The AGCS algorithm, supplemented by the ability to use search information to solve MMCOlex problems, will be referred to below as the global search algorithm for multistage solutions of a set of multiextremal optimization problems with nonlinear constraints (MAGCS).
 
@@
 
-For MAGCS, 50 subproblems (\ref{eq:30}) were solved for different values $\theta=\theta_1$ from (\ref{eq:31}), uniformly distributed in the interval $[0,1]$.
+For MAGCS, the problem (\ref{eq:43}) is considered as the MMCOlex problem, which includes 50 subproblems (\ref{eq:30}) with different values
+$\theta_1$ from (\ref{eq:31}), uniformly distributed in the interval $[0,1]$.
 The accuracy $\varepsilon=0.05$ from (\ref{eq:28}) and the reliability parameters $r_1=r_2=3.0$ from (\ref{eq:29}) were used.
@@
 
-All the subsequent numerical experiments focused on the solution of 100 two-criteria MCO problems, the efficiency criteria of which were set by the GKLS generator.
-For each MCO problem from the generated set, from 10 to 100 subproblems (\ref{eq:30}) were solved for different values $\theta=\theta_1$ from (\ref{eq:31}), uniformly distributed in the interval $[0,1]$.
-The obtained results of numerical experiments were averaged by the number of solved MCO problems.
+All the subsequent numerical experiments focused on the solution of 
+the MMCOlex problems, which includes 100 two-criteria MCO problems, the efficiency criteria of which were set by the GKLS generator.
+Each MCO problem from the generated set is considered as the MMCOlex problem, which includes from 10 to 100 subproblems (\ref{eq:30}) with different
+values $\theta_1$ from (\ref{eq:31}), uniformly distributed in the interval $[0,1]$. The obtained results of numerical experiments were averaged by the number of solved MCO problems.
 
-The solution of MCO problems was carried out before the stopping criterion of (\ref{eq:28}) was fulfilled, and the MCO problems were considered solved correctly if the distance of the decisons to the Pareto set did not exceed the given accuracy $\varepsilon$ from (\ref{eq:28}).
+The solution of MCOlex problems was carried out before the stopping criterion of (\ref{eq:28}) was fulfilled, and the MCOlex problems were considered solved correctly if the distance of the solutions to the Pareto set did not exceed the given accuracy $\varepsilon$ from (\ref{eq:28}).
 
@@
 
-The first column (SP) in the Table \ref{tab:2} contains the total number of MCO subproblems to be solved.
+The first column (SP) in the Table \ref{tab:2} contains the total number of MCOlex subproblems to be solved.
 The PI columns show the average number of iterations for solving one MCO problem.
-The SPI columns represent the average number of iterations for solve a single MCO subproblem.
-The SPS columns contain a fraction of MCO subproblems solved with the required accuracy.
+The SPI columns represent the average number of iterations for solve a single MCOlex subproblem.
+The SPS columns contain a fraction of MCOlex subproblems solved with the required accuracy.
 The HV and DU columns show the hypervolume (HV) and the distribution uniformity (DU) indices of the calculated approximation of the Pareto set, respectively.
@@
 The results presented in Table \ref{tab:2} demonstrate that the use of search information allows to obtain an estimate of the Pareto set after a substantially smaller total number of iterations of the global search (the PI columns).
-It should also be noted that as the number of solved subproblems increases, the MCO allows to build estimates of the Pareto set with better HV and DU indices.
+It should also be noted that as the number of solved subproblems increases, the MAGCS allows to build estimates of the Pareto set with better HV and DU indices.
 
 For a better visual representation, Fig.
-\ref{fig:4} shows the average number of iterations performed to solve a single MCO subproblem with and without search information.
-The presented graphs convincingly demonstrate a significant reduction in the number of global search iterations when solving the MCO subproblems using search information.
+\ref{fig:4} shows the average number of iterations performed to solve a single MCOlex subproblem with and without search information.
+The presented graphs convincingly demonstrate a significant reduction in the number of global search iterations when solving the MCOlex subproblems using search information.
 
@@
   \includegraphics[width=1\linewidth]{fig4}
-  \caption{Average number of iterations required to solve one MCO subproblem depending on the total number of subproblems solved}
+  \caption{Average number of iterations required to solve one MCOlex subproblem depending on the total number of subproblems solved}
   \label{fig:4}
@@
 Fig.
-\ref{fig:5} shows the reduction in the number of global search iterations when solving MCO subproblems using search information.
-The graphs presented show that when the number of solved MCO subproblems grows linearly, the number of iterations required to solve one subproblem is reduced almost linearly as well.
+\ref{fig:5} shows the reduction in the number of global search iterations when solving MCOlex subproblems using search information.
+The graphs presented show that when the number of solved MCOlex subproblems grows linearly, the number of iterations required to solve one subproblem is reduced almost linearly as well.
 
@@
   \includegraphics[width=1\linewidth]{fig5}
-  \caption{Reduction of the number of iterations performed to solve one MCO subproblem depending on the total number of subproblems solved}
+  \caption{Reduction of the number of iterations performed to solve one MCOlex subproblem depending on the total number of subproblems solved}
   \label{fig:5}
@@
 
-It should be noted that the number of iterations of the global search is also reduced in the course of sequential solution of subproblems of the same MCO problem.
-
 It should be noted that the number of global search iterations is also reduced in the course of sequential solution of the subproblems of the same MCO problem.
-Table \ref{tab:3} illustrates this trend showing the average number of iterations required to find a solution to one MCO subproblem, with subdividing the sequence of solved subproblems in groups of 10 MCO subproblems.
-The first row (SP) contains the MCO subproblem numbers belonging to the individual subproblem groups.
+Table \ref{tab:3} illustrates this trend showing the average number of iterations required to find a solution to one MCOlex subproblem, with subdividing the sequence of solved subproblems in groups of 10 MCOlex subproblems.
+The first row (SP) contains the MCOlex subproblem numbers belonging to the individual subproblem groups.
 The PI rows show the total number of iterations of the global search performed for all previous problems.
-The rows labeled as ``PI new'' show the number of iterations performed to solve the next group of MCO subproblems with the numbers of the problems given in each column heading.
-The last row labeled as ``FPI new'' shows the proportion of new iterations of global search performed to solve the next group of MCO subproblems.
+The rows labeled as ``PI new'' show the number of iterations performed to solve the next group of MCOlex subproblems with the numbers of the problems given in each column heading.
+The last row labeled as ``FPI new'' shows the proportion of new iterations of global search performed to solve the next group of MCOlex subproblems.
 
@@
 \ref{fig:6} shows the values of the achieved reduction in the number of additional iterations required to solve the next set of subproblems (the last row of Table \ref{tab:3}).
-The above data demonstrates that solving the next group of MCO subproblems requires about $6\%$ of new global search iterations in addition to the available search information.
+The above data demonstrates that solving the next group of MCOlex subproblems requires about $6\%$ of new global search iterations in addition to the available search information.
 
@@
   \includegraphics[width=1\linewidth]{fig6}
-  \caption{Proportion of new global search iterations for solving the next group of MCO subproblems}
+  \caption{Proportion of new global search iterations for solving the next group of MCOlex subproblems}
   \label{fig:6}
@@
 \ref{fig:4}--\ref{fig:6}.
-These data demonstrate that the experiments carried out to solve more complex MCO problems confirm the conclusions made earlier -- for example, solving the next group of MCO subproblems requires, in addition to the available search information, from $0.5\%$ to $14.2\%$ more global search iterations (Table \ref{tab:5}, last row).
+These data demonstrate that the experiments carried out to solve more complex MCO problems confirm the conclusions made earlier -- for example, solving the next group of MCOlex subproblems requires, in addition to the available search information, from $0.5\%$ to $14.2\%$ more global search iterations (Table \ref{tab:5}, last row).
 
@@
 A key feature of this class of problems is the ability to alter the ranking of efficiency criteria in terms of importance during calculations, which in turn necessitates a multistage solution for MCOlex problems.
+This new type of problems is referred to as MMCOlex problems.
 
-The computational complexity of addressing the formulated new class of MCOlex problems can be overcome by solving a sequence of global optimization problems with nonlinear constraints using efficient information-statistical methods of global optimization.
-A core element in this approach is the ability to use all the search information obtained in the course of solving the multistage MCOlex problems.
+The computational complexity of addressing the formulated MMCOlex problems can be overcome by solving a sequence of global optimization problems with nonlinear constraints using efficient information-statistical methods of global optimization.
+A core element in this approach is the ability to use all the search information obtained in the course of solving the MMCOlex problems.
 With each new stage of the solution, this search information allows incorporating the previously calculated values of efficiency criteria into the values of the next scalar problem of multiextremal optimization.
@@
 
-The results of numerical experiments demonstrate that this approach can sufficiently reduce the computational complexity of multistage MCOlex problem-solving.
+The results of numerical experiments demonstrate that this approach can sufficiently reduce the computational complexity of MMCOlex problem-solving.
 

### 2020-10-07 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
comments in russian

@@
 \usepackage{cite}
+
+\usepackage[utf8x]{inputenc}
+\usepackage[english,russian]{babel}
+
+\usepackage{xcolor}
 
@@
 
-A note worthy approach to solving MCO problems is lexicographic optimization methods.
+A noteworthy approach to solving MCO problems is lexicographic optimization methods.
 Such methods involve ranking the criteria by the degree of their importance and using the obtained ranking for subsequent optimization \cite{c3}.
@@
 Problem (\ref{eq:1}) is considered in the context of the most complex decision-making problems, in which the efficiency criteria $f_i(y)$, $1 \leq i \leq s$, can be significantly multiextremal, and the procedures for calculating the values of the efficiency criteria at points of the search domain $y \in D$ can be computationally expensive.
-The criteria $f_i (y)$, $1 \leq i \leq s$ are also assumed to satisfy the Lipschitz condition
+The criteria $f_i (y)$, $1 \leq i \leq s$, are also assumed to satisfy the Lipschitz condition
 \begin{equation}\label{eq:3}
@@
 \end{equation}
-%new text begin
+{\color{red} 
 where $F_t, t>1,$ is a MCOlex problem. The size of the set $\mathbb{F}$ can be changed 
@@
 In general, this approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex)\footnote{From a general point of view, the MCOlex problem can be considered as a special case of the MMCOlex problem for $t=1$.}.
-%new text end
+}
 
@@
 	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function at these points), which significantly increases the complexity and computational cost of the global optimization methods.
-Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47}, the Peano space-filling curves \cite{c17,c23}.
+Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47,Grishagin2016_1,Grishagin2019}, the Peano space-filling curves \cite{c17,c23}.
 	\item When solving multiple problems of multistage lexicographic multicriteria optimization, global search methods can reduce the complexity of calculations by using search information obtained during the calculations.
@@
 
-Dimension reduction by the Peano curves is widely used in the development of efficient global search algorithms in the framework of the information-statistical theory of multiextremal optimization \cite{c17,c18,c23,c26,c27}.
+Dimension reduction by the Peano curves is widely used in the development of efficient global search algorithms in the framework of the information-statistical theory of multiextremal optimization \cite{c17,c18,c23,c26,c27,Grishagin2016_2}.
 Numerical methods for constructing approximations of the Peano curves (evolvents) with a given accuracy are considered in \cite{c17,c18}.
@@
 \end{equation}
-(when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index allows determining the integrated function for the problem (\ref{eq:16}) (see Fig. \ref{fig:1})
+(when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index allows determining the integrated function for the problem (\ref{eq:16}) (see Fig. \ref{fig:1},
+{\color{red}  the example taken from \cite{Book2013}})
 \begin{equation}\label{eq:19}
@@
   \includegraphics[width=\linewidth]{fig1}
-  \caption{An example of the integrated function for the problem (\ref{eq:16}) using a sequential constraint calculation scheme: a) the function $F(x)$ from (\ref{eq:19}), b) the function $\Phi(x)$ from (\ref{eq:20})}
+  \caption{An example of the integrated function for the problem (\ref{eq:16}) using a sequential constraint calculation scheme: a) {\color{red} initial problem functions} , b) the function $F(x)$ from (\ref{eq:19})}
   \label{fig:1}
@@
 
-Various modifications of this algorithm and the corresponding theory of convergence are presented in \cite{c18}. For instance, the following theorem is valid.
+Various modifications of this algorithm and the corresponding theory of convergence are presented in \cite{c18}.
+{\color{red}
+В частности, одна из модификаций связана с использованием информации о локальном поведении функций задачи \cite{Sergeyev2003}.
+Условия сходимости алгоритма определяются следующей теоремой из \cite{c18}.
+}
 
@@
 Generally, the solution to such problems should start from the very beginning; and multiple solutions to such global optimization problems may require extensive calculations.
-However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}.
-%new text begin
+However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_j (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}.
+{\color{red} 
 This possibility of reusing search information can be briefly presented as follows.
-%\begin{itemize}
-%\item 
-%При минимизации очередного критерия $f_l(y(x)), 1\leq l\leq s,$ получаемая поисковая информация накапливается в форме множества $A_k$ из (\ref{eq:21}) (следует отметить, что в $A_k$ наряду со значениями $z_i, 1\leq i\leq k,$ минимизируемого критерия $f_l(x)$ запоминаются и вычисленные значения критериев $f_i, 1\leq i\leq k$.
-%\item 
-%При переходе к минимизации следующего критерия $f_{l+1}(y(x))$ значения этого критерия $z_i, 1\leq i\leq k,$ во всех точках $x_i, 1\leq i\leq k,$ ранее выполненных итераций глобального поиска могут получены при помощи быстро выполнимых вычислений на основе запомненных значений критериев $f_i, 1\leq i\leq k$ в поисковой информации $A_k$
-%\begin{equation}\label{new_equation}
-%z'_i=Z(f_i), 1\leq i\leq k,
-%\end{equation}
-%где процедура вычисления значений $z'_i, 1\leq i\leq k$ определяется правилами (\ref{eq:9}), (\ref{eq:18})--(\ref{eq:20}) без повторных вычислениях значений критериев $f_l(y(x)), 1\leq l\leq s$.
-%\end{itemize}
-%new text end
+\begin{itemize}
+\item 
+При минимизации очередного критерия $f_j(y(x)), 1\leq j\leq s,$ получаемая поисковая информация накапливается в форме множества $A_k$ из (\ref{eq:21}) (следует отметить, что в $A_k$ наряду со значениями $z_i, 1\leq i\leq k,$ минимизируемого критерия $f_j(y(x))$ запоминаются и вычисленные значения векторного критерия $f(y(x_i)), 1\leq i\leq k$.
+\item 
+При переходе к минимизации следующего критерия $f_{j+1}(y(x))$ значения этого критерия $z_i, 1\leq i\leq k,$ во всех точках $x_i, 1\leq i\leq k,$ ранее выполненных итераций глобального поиска могут получены при помощи быстро выполнимых вычислений на основе запомненных значений векторного критерия $f(y(x_i)), 1\leq i\leq k,$ в поисковой информации $A_k$
+\begin{equation}\label{new_equation}
+z'_i=Z(f(y(x_i))), 1\leq i\leq k,
+\end{equation}
+где процедура вычисления значений $z'_i, 1\leq i\leq k$ определяется правилами (\ref{eq:9}), (\ref{eq:18})--(\ref{eq:20}), без повторных вычислениях значений критериев $f_j(y(x)), 1\leq l\leq s$.
+\end{itemize}
+}
 This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of numerical experiments.
@@
 \end{equation}
-%new text begin
-%При этом, как и ранее, при любых изменениях оценок  $f_j^*, \delta_j, g_j^k, 1\leq j < s, 1 \leq i \leq k,$ решение задачи (\ref{eq:30}) может быть продолжено с использованием всей имеющейся поисковой информации после быстро выполняемых вычислений новых значений $z'_i,1\leq i \leq k,$ в соответствии с процедурой (\ref{new_equation}).
-%new text end
+{\color{red} 
+При этом, как и ранее, при любых изменениях оценок  $f_j^*, \delta_j, g_j^k, 1\leq j < s, 1 \leq i \leq k,$ решение задачи (\ref{eq:30}) может быть продолжено с использованием всей имеющейся поисковой информации после быстро выполняемых вычислений новых значений $z'_i,1\leq i \leq k,$ в соответствии с процедурой (\ref{new_equation}).
+}
 In particular, a similar one-step scheme (\ref{eq:30})--(\ref{eq:31}) can be applied through the use of AGCS.
@@
 \begin{theorem}
-Let the next minimized function $\psi(x)$ of the computational scheme (\ref{eq:9}) differs from the previous minimized function $\phi(x)$ by some limited value $\delta$, i.e.
+Let the next minimized function $\psi(x)$ of the computational scheme (\ref{eq:9}) differ from the previous minimized function $\phi(x)$ by some limited value $\delta$, i.e.
 \begin{equation}\label{eq:32}
@@
 \begin{equation}\label{eq:36}
-L=\max{L_i}, 1 \leq i \leq s, L_i\; \text{from}\; (\ref{eq:3}),
+L=\max{L_i}, 1 \leq i \leq s, \; L_i\; \text{from}\; (\ref{eq:3}),
+\end{equation}
+\begin{equation}\label{eq:beta}
+\beta =  \frac{2r\delta} {L\varepsilon(r-1)},
 \end{equation}
@@
 A key feature of this class of problems is the ability to alter the ranking of efficiency criteria in terms of importance during calculations, which in turn necessitates a multistage solution for MCOlex problems.
-This new type of problems is referred to as MMCOlex problems.
+{\color{red} This new type of problems is referred to as MMCOlex problems.}
 
@@
 Optim., \textbf{71}(1), 73--90 (2018).
-doi:10.1007/s10898-018-0624-3.
+DOI:10.1007/s10898-018-0624-3.
 \bibitem{c49} Mart\'l, L., Garc\'la, J., Berlanga, A., Molina, J.M.: A stopping criterion for multi-objective optimization evolutionary algorithms. Inf. Sci. \textbf{367--368}, 700--718 (2016)
-\bibitem{c50} Dilettoso, E., Rizzo, S.A., Salerno, N.: A weakly pareto compliant quality indicator. Math. Comput. Appl. \textbf{22}(1), 25 (2017)
+\bibitem{c50} Dilettoso, E., Rizzo, S.A., Salerno, N.: A weakly Pareto compliant quality indicator. Math. Comput. Appl. \textbf{22}(1), 25 (2017)
 \bibitem{c51} Wu, J., Azarm, S.: Metrics for quality assessment of a multiobjective design optimization solution set. J. Mech. Des. \textbf{123}(1), 18--25 (2001)
-\bibitem{c52} Audet, C., Bigeon, J., Cartier, D., Digabel, S.Le, Salomon, L.: Performance indicators in multiobjective optimization. Technical Report G-2018-90, Les cahiers du GERAD. (2018)
+\bibitem{c52} Audet, C., Bigeon, J., Cartier, D., Le Digabel, S., Salomon, L.: Performance indicators in multiobjective optimization. Technical Report G-2018-90, Les cahiers du GERAD. (2018)
 \bibitem{c53} Huband, S., Hingston, P., Barone, L., While, L.: A review of multiobjective test problems and a scalable test problem toolkit. IEEE T. Evolut. Comput. \textbf{10}(5), 477--506 (2006)
@@
 \textbf{318}, 298--311 (2018)
-\bibitem{c59} Lai, L., Fiaschi, L., Cococcioni, M.: Solving mixed pareto-lexicographic multi-objective optimization problems: the case of priority chains.
+\bibitem{c59} Lai, L., Fiaschi, L., Cococcioni, M.: Solving mixed Pareto-lexicographic multi-objective optimization problems: the case of priority chains.
 Swarm Evol.
@@
 \textbf{55}, 100687 (2020) DOI: 10.1016/j.swevo.2020.100687
+
+\bibitem{Book2013} Strongin, R.G., Gergel, V.P., Grishagin, V.A., Barkalov, K.A.: Parallel computing in global optimization problems. MSU Press, Moscow. (2013, in Russian)
+
+\bibitem{Sergeyev2003} Sergeyev, Y.D., Pugliese, P., Famularo, D.: Index information algorithm with local tuning for solving multidimensional global optimization problems with multiextremal constraints.
+Math.
+Program., Ser.
+A \textbf{96}, 489--512 (2003)
+
+\bibitem{Grishagin2016_1} Grishagin, V.A., Israfilov, R.A.: Global search acceleration in the nested optimization scheme. In: AIP Conf. Proc., \textbf{1738}, 400010 (2016). DOI: 10.1063/1.4952198
+
+\bibitem{Grishagin2016_2} Grishagin, V.A., Israfilov, R.A., Sergeyev, Y.D.: Comparative efficiency of dimensionality reduction schemes in global optimization.
+In: AIP Conf.
+Proc., \textbf{1776}, 060011 (2016).
+DOI: 10.1063/1.4965345
+
+\bibitem{Grishagin2019} Gergel V.P., Grishagin V.A., Israfilov R.A.
+Adaptive dimensionality reduction in multiobjective optimization with multiextremal criteria.
+Lecture Notes in Computer Science, \textbf{11331}, 129--140 (2019)
 

### 2020-10-07 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
add one bracket

@@
 
-\usepackage{hyperref}
+%\usepackage{hyperref}
 \usepackage{cite}
@@
 However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_j (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}.
-{\color{red} 
-This possibility of reusing search information can be briefly presented as follows.
+{\color{red} This possibility of reusing search information can be briefly presented as follows.
 \begin{itemize}
 \item 
-При минимизации очередного критерия $f_j(y(x)), 1\leq j\leq s,$ получаемая поисковая информация накапливается в форме множества $A_k$ из (\ref{eq:21}) (следует отметить, что в $A_k$ наряду со значениями $z_i, 1\leq i\leq k,$ минимизируемого критерия $f_j(y(x))$ запоминаются и вычисленные значения векторного критерия $f(y(x_i)), 1\leq i\leq k$.
+При минимизации очередного критерия $f_j(y(x)), 1\leq j\leq s,$ получаемая поисковая информация накапливается в форме множества $A_k$ из (\ref{eq:21}) (следует отметить, что в $A_k$ наряду со значениями $z_i, 1\leq i\leq k,$ минимизируемого критерия $f_j(y(x))$ запоминаются и вычисленные значения векторного критерия $f(y(x_i)), 1\leq i\leq k$).
 \item 
-При переходе к минимизации следующего критерия $f_{j+1}(y(x))$ значения этого критерия $z_i, 1\leq i\leq k,$ во всех точках $x_i, 1\leq i\leq k,$ ранее выполненных итераций глобального поиска могут получены при помощи быстро выполнимых вычислений на основе запомненных значений векторного критерия $f(y(x_i)), 1\leq i\leq k,$ в поисковой информации $A_k$
+При переходе к минимизации следующего критерия $f_{j+1}(y(x))$ значения этого критерия $z'_i, 1\leq i\leq k,$ во всех точках $x_i, 1\leq i\leq k,$ ранее выполненных итераций глобального поиска могут получены при помощи быстро выполнимых вычислений на основе запомненных значений векторного критерия $f(y(x_i)), 1\leq i\leq k,$ в поисковой информации $A_k$
 \begin{equation}\label{new_equation}

### 2020-10-14 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
after translation of comments

--- 2020-10-07 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
@@
 {\color{red} 
-where $F_t, t>1,$ is a MCOlex problem. The size of the set $\mathbb{F}$ can be changed 
+where $F_t, \; t>1,$ is a MCOlex problem. The size of the set $\mathbb{F}$ can be changed 
 dynamically by adding new required MCOlex problems or removing unnecessary ones. 
@@
 (when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index allows determining the integrated function for the problem (\ref{eq:16}) (see Fig. \ref{fig:1},
-{\color{red}  the example taken from \cite{Book2013}})
+{\color{red}  the example was taken from \cite{Book2013}})
 \begin{equation}\label{eq:19}
@@
   \includegraphics[width=\linewidth]{fig1}
-  \caption{An example of the integrated function for the problem (\ref{eq:16}) using a sequential constraint calculation scheme: a) {\color{red} initial problem functions} , b) the function $F(x)$ from (\ref{eq:19})}
+  \caption{An example of the integrated function for the problem (\ref{eq:16}) using a sequential constraint calculation scheme: a) {\color{red} the initial problem functions}, b) the function $F(x)$ from (\ref{eq:19})}
   \label{fig:1}
@@
 {\color{red}
-В частности, одна из модификаций связана с использованием информации о локальном поведении функций задачи \cite{Sergeyev2003}.
-Условия сходимости алгоритма определяются следующей теоремой из \cite{c18}.
+In particular, one modification of the algorithm involves using information about local behavior of the problem function  \cite{Sergeyev2003}.
+The AGCS convergence conditions are defined by the following theorem from  \cite{c18}.
 }
@@
 However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_j (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}.
-{\color{red} 
-This possibility of reusing search information can be briefly presented as follows.
+{\color{red} The search information can be reused by using the algorithm briefly described below.
 \begin{itemize}
 \item 
-При минимизации очередного критерия $f_j(y(x)), 1\leq j\leq s,$ получаемая поисковая информация накапливается в форме множества $A_k$ из (\ref{eq:21}) (следует отметить, что в $A_k$ наряду со значениями $z_i, 1\leq i\leq k,$ минимизируемого критерия $f_j(y(x))$ запоминаются и вычисленные значения векторного критерия $f(y(x_i)), 1\leq i\leq k$.
+When the criterion $f_j(y(x)), 1\leq j\leq s,$ is minimized, the resulting search information is accumulated as a set $A_k$ from (\ref{eq:21}) (note that along with the values $z_i, 1\leq i\leq k,$ of the minimized criterion $f_j(y(x))$, $A_k$ also accumulates the calculated values of the vector criterion $f(y(x_i)), 1\leq i\leq k$).
 \item 
-При переходе к минимизации следующего критерия $f_{j+1}(y(x))$ значения этого критерия $z_i, 1\leq i\leq k,$ во всех точках $x_i, 1\leq i\leq k,$ ранее выполненных итераций глобального поиска могут получены при помощи быстро выполнимых вычислений на основе запомненных значений векторного критерия $f(y(x_i)), 1\leq i\leq k,$ в поисковой информации $A_k$
+When the minimization of the next criterion $f_{j+1}(y(x))$ begins, the values of this criterion $z'_i, 1\leq i\leq k,$ at all the points  $x_i, 1\leq i\leq k,$ of the previously completed iterations of the global search can be obtained via quick calculations based on the stored values of the vector criterion $f(y(x_i)), 1\leq i\leq k,$ in the search information $A_k$
 \begin{equation}\label{new_equation}
@@
 \end{equation}
-где процедура вычисления значений $z'_i, 1\leq i\leq k$ определяется правилами (\ref{eq:9}), (\ref{eq:18})--(\ref{eq:20}), без повторных вычислениях значений критериев $f_j(y(x)), 1\leq l\leq s$.
+where calculation procedure of the values $z'_i, 1\leq i\leq k$ is determined by the rules (\ref{eq:9}), (\ref{eq:18})--(\ref{eq:20}), without any need to repeat calculations of criteria values $f_j(y(x)), 1\leq l\leq s$.
 \end{itemize}
@@
 \begin{equation}\label{eq:30}
-P_{\delta lex} (f,\delta,D)=Arg \min_{y \in D} {f_s(y)}, f_j(y)\leq f_j^* + \delta_j, 1 \leq j < s,
+P_{\delta lex} (f,\delta,D)=\arg \min_{y \in D} { \{ f_s(y), f_j(y)\leq f_j^* + \delta_j, 1 \leq j < s\} },
 \end{equation}
@@
 {\color{red} 
-При этом, как и ранее, при любых изменениях оценок  $f_j^*, \delta_j, g_j^k, 1\leq j < s, 1 \leq i \leq k,$ решение задачи (\ref{eq:30}) может быть продолжено с использованием всей имеющейся поисковой информации после быстро выполняемых вычислений новых значений $z'_i,1\leq i \leq k,$ в соответствии с процедурой (\ref{new_equation}).
+Considering this and previous examples, when any changes in estimates $f_j^*, \delta_j,$ $g_j^k, 1\leq j < s, 1 \leq i \leq k,$ occur, the solution of the problem (\ref{eq:30}) can be continued using all the available search information after the new values $z'_i,1\leq i \leq k,$ are promptly calculated following the procedure (\ref{new_equation}).
 }
@@
 \begin{equation}\label{eq:35}
-m=\max{r_l H_l^k}, 1 \leq l \leq M^k, k>k_0,
+m=\max{\{r_l H_l^k, 1 \leq l \leq M^k \} }, k>k_0,
 \end{equation}
 \begin{equation}\label{eq:36}
-L=\max{L_i}, 1 \leq i \leq s, \; L_i\; \text{from}\; (\ref{eq:3}),
+L=\max{\{L_i, 1 \leq i \leq s \}}, \; L_i\; \text{from}\; (\ref{eq:3}),
 \end{equation}
@@
 \begin{equation}\label{eq:37}
-\psi(x^*)=\min{\psi(x)},x \in [0,1],
+\psi(x^*)=\min{\{\psi(x), x \in [0,1]\}},
 \end{equation}
 \begin{equation}\label{eq:38}
-x_k^* = arg \min{\phi(x_k)}, 1 \leq i \leq k.
+x_k^* = arg \min{\{\phi(x_k), 1 \leq i \leq k\}}.
 \end{equation}

### 2020-10-14 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
correct agreement number

@@
 
-\title{Computationally Efficient Approach for Solving Lexicographic Multicriteria Optimization Problems\thanks{The work was supported by the Ministry of Science and Higher Education of the Russian Federation, agreement No 13.1902.21.0026.
+\title{Computationally Efficient Approach for Solving Lexicographic Multicriteria Optimization Problems\thanks{The work was supported by the Ministry of Science and Higher Education of the Russian Federation, agreement No 075-15-2020-808.
 %``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''.

### 2020-10-15 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
fix some misprints

@@
 it must be dense only in the vicinity of globally optimal points.
-	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- one of the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}).
+	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- one of the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:3}).
 	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function at these points), which significantly increases the complexity and computational cost of the global optimization methods.
@@
 \end{equation}
-where the constants $H_i$ are determined by the ratio $H_i=2L_i\sqrt{N+3}$,  $1 \leq i \leq m$, the values $L_i$ are the Lipschitz constants from (\ref{eq:2}), and $N$ is the dimension of the optimization problem (\ref{eq:1}).
+where the constants $H_i$ are determined by the ratio $H_i=2L_i\sqrt{N+3}$,  $1 \leq i \leq s$, the values $L_i$ are the Lipschitz constants from (\ref{eq:3}), and $N$ is the dimension of the optimization problem (\ref{eq:1}).
 
@@
 \begin{equation} \label{eq:15}
-x_1 < x_2 < \dots < x_k	
+x_1 < x_2 < \dots < x_k,	
 \end{equation}
@@
 
-To provide uniform calculations, a sequential scheme is applied to obtain values the constraints $g_i(y(x)) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$.
+To provide uniform calculations, a sequential scheme is applied to obtain values the constraints $g_i(x) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$.
 According to this scheme, the constraint values are calculated strictly in their order, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme).
@@
 \begin{equation}\label{eq:18}
-\nu(x) : g_\nu(x)>0, g_j(x)\leq 0, 1 \leq j \leq \nu - 1, 1 \leq \nu = \nu(x) \leq m
+\nu(x) : g_\nu(x)>0, g_j(x)\leq 0, 1 \leq j \leq \nu - 1, 1 \leq \nu = \nu(x) \leq m.
 \end{equation}
-(when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index allows determining the integrated function for the problem (\ref{eq:16}) (see Fig. \ref{fig:1},
-{\color{red}  the example was taken from \cite{Book2013}})
+When all the constraints are satisfied, $\nu=m+1$ is assumed. Introducing the index allows determining the integrated function for the problem (\ref{eq:16})
 \begin{equation}\label{eq:19}
-F(x)=g_\nu(x), \nu=\nu(x),
+F(x)=g_\nu(x), \; \nu=\nu(x),
 \end{equation}
-defined and computable everywhere in $[0,1]$.
+defined and computable everywhere in $[0,1]$ (see Fig.~\ref{fig:1},
+{\color{red}  the example was taken from \cite{Book2013}}).
 Its value at the point $x\in [0,1]$ is either the value of the left side of the constraint violated at this point (the case when $\nu \leq m$), or the value of the minimized function (the case when $\nu = m + 1$).

### 2020-10-16 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
remove red color

@@
 
-\usepackage{xcolor}
+%\usepackage{xcolor}
 
@@
 \begin{equation}\label{eq:8}
-D_1=Arg \min_{y \in D}{f_1(y)},\dots, D_i=Arg \min_{y \in D_{i-1}}{f_i(y)},1 \leq i \leq s,D_0=D,
+D_1=\arg \min_{y \in D}{f_1(y)},\dots, D_i=\arg \min_{y \in D_{i-1}}{f_i(y)},1 \leq i \leq s,D_0=D,
 \end{equation}
-where $Arg$ is the set of all decisions that achieve the minimum value of the optimized criterion.
+where $arg$ is the set of all decisions that achieve the minimum value of the optimized criterion.
 It should be noted that the sequence of steps in (\ref{eq:8}) may not be fully completed if at a certain stage $l$, $1 \leq l \leq s$, the set $D_l$ degenerates and contains only a single decision $y \in D$.
@@
 f^*_i =\min_{y \in D}{f_i(y)},  f_j(y) \leq f^*_j + \delta_j, 1 \leq j < i, 1 < i \leq s, \\
-P_{lex}^\delta(f,\delta,D) = Arg \min_{y \in D}{f_s(y)},f_j(y) \leq f^*_j + \delta_j, 1 \leq j < s.
+P_{lex}^\delta(f,\delta,D) = \arg \min_{y \in D}{f_s(y)},f_j(y) \leq f^*_j + \delta_j, 1 \leq j < s.
 \end{matrix}
@@
 \end{equation}
-{\color{red} 
-where $F_t, \; t>1,$ is a MCOlex problem. The size of the set $\mathbb{F}$ can be changed 
+where $F_i, \; 1\leq i \leq t,$ is a MCOlex problem. The size of the set $\mathbb{F}$ can be changed 
 dynamically by adding new required MCOlex problems or removing unnecessary ones. 
@@
 In general, this approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex)\footnote{From a general point of view, the MCOlex problem can be considered as a special case of the MMCOlex problem for $t=1$.}.
-}
 
@@
 \end{equation}
-defined and computable everywhere in $[0,1]$ (see Fig.~\ref{fig:1},
-{\color{red}  the example was taken from \cite{Book2013}}).
+defined and computable everywhere in $[0,1]$ (see Fig.~\ref{fig:1}, the example was taken from \cite{Book2013}).
 Its value at the point $x\in [0,1]$ is either the value of the left side of the constraint violated at this point (the case when $\nu \leq m$), or the value of the minimized function (the case when $\nu = m + 1$).
@@
   \includegraphics[width=\linewidth]{fig1}
-  \caption{An example of the integrated function for the problem (\ref{eq:16}) using a sequential constraint calculation scheme: a) {\color{red} the initial problem functions}, b) the function $F(x)$ from (\ref{eq:19})}
+  \caption{An example of the integrated function for the problem (\ref{eq:16}) using a sequential constraint calculation scheme: a) the initial problem functions, b) the function $F(x)$ from (\ref{eq:19})}
   \label{fig:1}
@@
 Various modifications of this algorithm and the corresponding theory of convergence are presented in \cite{c18}.
-{\color{red}
 In particular, one modification of the algorithm involves using information about local behavior of the problem function  \cite{Sergeyev2003}.
+
 The AGCS convergence conditions are defined by the following theorem from  \cite{c18}.
-}
 
@@
 However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_j (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}.
-{\color{red} The search information can be reused by using the algorithm briefly described below.
+The search information can be reused by using the algorithm briefly described below.
 \begin{itemize}
@@
 \end{itemize}
-}
+
 This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of numerical experiments.
@@
 \end{equation}
-{\color{red} 
 Considering this and previous examples, when any changes in estimates $f_j^*, \delta_j,$ $g_j^k, 1\leq j < s, 1 \leq i \leq k,$ occur, the solution of the problem (\ref{eq:30}) can be continued using all the available search information after the new values $z'_i,1\leq i \leq k,$ are promptly calculated following the procedure (\ref{new_equation}).
-}
 In particular, a similar one-step scheme (\ref{eq:30})--(\ref{eq:31}) can be applied through the use of AGCS.
@@
 \begin{equation}\label{eq:38}
-x_k^* = arg \min{\{\phi(x_k), 1 \leq i \leq k\}}.
+x_k^* = \arg \min{\{\phi(x_k), 1 \leq i \leq k\}}.
 \end{equation}
@@
 A key feature of this class of problems is the ability to alter the ranking of efficiency criteria in terms of importance during calculations, which in turn necessitates a multistage solution for MCOlex problems.
-{\color{red} This new type of problems is referred to as MMCOlex problems.}
+This new type of problems is referred to as MMCOlex problems.f
 

### 2020-10-16 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
sorted the list of references

@@
 
-The initial theoretical analysis of global search properties when using search information was performed in \cite{c48}.
+The initial theoretical analysis of global search properties when using search information was performed in \cite{c31}.
 The obtained theoretical statements can be reformulated as applied to lexicographic multicriteria optimization problems.
@@
 \bibitem{c9} Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications. Surv. Math. Ind. \textbf{11}, 1--42 (2005)
+\bibitem{c14} Eichfelder, G.: Scalarizations for adaptively solving multi-objective optimization problems. Comput. Optim. Appl. \textbf{44}, 249--273 (2009)
 \bibitem{c10} Branke, J., Deb, K., Miettinen, K., Slowinski, R., (eds.): Multi-Objective Optimization--Interactive and Evolutionary Approaches. Springer, Berlin. (2008)
@@
 \bibitem{c13} Tan, K.C., Khor, E.F., Lee, T.H.: Multi-objective Evolutionary Algorithms and Applications. Springer-Verlag, London. (2005)
-\bibitem{c14} Eichfelder, G.: Scalarizations for adaptively solving multi-objective optimization problems. Comput. Optim. Appl. \textbf{44}, 249--273 (2009)
-\bibitem{c15} Jones, D.R.: A taxonomy of global optimization methods based on response surfaces. J. Glob. Optim. \textbf{21}, 345--383 (2001)
-\bibitem{c16} Voutchkov, I., Keane, A.: Multi-objective optimization using surrogates. Comput. Intell. Optim. Adapt. Learn. Optim. \textbf{7}, 155--175 (2010)
-\bibitem{c17} Strongin, R.G.: Numerical Methods in Multiextremal Problems: Information-statistical Algorithms. Nauka, Moscow. (1978, in Russian) 
-\bibitem{c18} Strongin, R.G., Sergeyev, Y.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht. (2nd ed. 2013, 3rd ed. 2014).
-\bibitem{c19} T\"orn, A., {\v Z}ilinskas, A.: Global Optimization. Lecture Notes in Computer Science, 350. Springer-Verlag, Berlin. (1989)
-\bibitem{c20} Horst, R., Tuy, H.: Global Optimization: Deterministic Approaches. Springer-Verlag, Berlin. (1990)
-\bibitem{c21} Zhigljavsky, A.A.: Theory of Global Random Search. Kluwer Academic Publishers, Dordrecht. (1991)
-\bibitem{c22} Pint\'er, J.D.: Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications). Kluwer Academic Publishers, Dortrecht. (1996)
-\bibitem{c23} Sergeyev, Y.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting Space-filling Curves. Springer. (2013)
-\bibitem{c24} Locatelli, M., Schoen, F.: Global Optimization: Theory, Algorithms, and Applications. SIAM. (2013)
-\bibitem{c25} Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. Princeton University Press. (2016)
-\bibitem{c26} Lera, D., Sergeyev, Y.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and Holder constants.
-Commun.
-Nonlinear Sci.
-\textbf{23}(1-3), 328--342 (2015)
-\bibitem{c27} Gergel, V.: A unified approach to use of coprocessors of various types for solving global optimization problems. 2nd Int. Conf. on MACSI. 13--18 (2015). DOI: 10.1109/MCSI.2015.18
-\bibitem{c28} Arora, R.K.: Optimization: Algorithms and Applications. CRC Press. (2015)
-\bibitem{c29} Bazaraa, M.S., Sherali, H.D., Shetty, C.M.: Nonlinear Programming: Theory and Algorithms. John Wiley and Sons. (2006)
-\bibitem{c30} Gergel, V.P., Kozinov, E.A.: Accelerating multicriterial optimization by the intensive exploitation of accumulated search data.
-In: 	AIP Conf.
-Proc., \textbf{1776}, 090003 (2016).
-DOI: 10.1063/1.4965367
-\bibitem{c31} Gergel, V.P., Kozinov, E.A.: Efficient multicriterial optimization based on intensive reuse of search information.
-J.
-Glob.
-Optim.
-\textbf{71}(1), 73--90 (2018).
-DOI: 10.1007/s10898-018-0624-3
-\bibitem{c32} Sysoyev, A., Barkalov, K., Gergel, V.: Globalizer: A novel supercomputer software system for solving time-consuming global optimization problems.
-Numer.
-Algebra, Control.
-Optim.
-\textbf{8}(1), 47--62 (2018)
-\bibitem{c33} Kvasov, D.E., Sergeyev, Y.D.: Deterministic approaches for solving practical black-box global optimization problems. Adv. Eng. Softw. \textbf{80}, 58--66 (2015)
-\bibitem{c34} Modorskii, V., Gaynutdinova, D., Gergel, V., Barkalov, K.: Optimization in design of scientfic products for purposes of cavitation problems. AIP Conf. Proc. \textbf{1738}, 400013 (2016)
-\bibitem{c35} Evtushenko, Yu.G., Posypkin, M.A.: A deterministic algorithm for global multi-objective optimization. Optim. Methods Softw., \textbf{29}(5), 1005--1019 (2014)
-\bibitem{c36} {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Adaptation of a one-step worst-case optimal univariate algorithm of bi-objective Lipschitz optimization to multidimensional problems.
-Commun.
-Nonlinear Sci.
-Numer.
-Simul., \textbf{21}(1-3), 89--98 (2015)
-\bibitem{c37} Gaviano, M., Kvasov, D.E, Lera, D., Sergeyev, Y.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
-ACM Trans.
-Math.
-Softw.
-\textbf{29}(4), 469--480 (2003)
 
-\bibitem{c38} Zarepisheh, M., Khorram, E.: On the transformation of lexicographic nonlinear multiobjective  programs to single objective programs. Math. Methods Oper. Res. \textbf{74}, 217--231 (2011)
-\bibitem{c39} Haimes, Y.Y., Lasdon, L., Wismer, D.: On a bicriterion formulation of the problem of integrated systems identification and system optimization.
-IEEE Trans.
-Syst.
-Man Cybern.
-\textbf{1}, 296--297 (1971)
-\bibitem{c40} Chankong, V., Haimes, Y.Y.: Multiobjective Decision Making: Theory and Methodology. North-Holland Series in System Science and Engineering. New York (NY): Elsevier. (1983)
-\bibitem{c41} Ehrgott, M., Ruzika, S.: Improved $\varepsilon$-constraint method for multiobjective programming. J. Optim. Theory Appl. \textbf{138}, 375--396 (2008)
-\bibitem{c42} Rastegar, N., Khorram, E.: Relaxation of constraints in lexicographic multiobjective programming problems.
-A Journal of Mathematical Programming and Operations Research, \textbf{64}, 2111--2129 (2015)
-\bibitem{c43} Castro-Gutierrez, J., Landa-Silva, D., P\'erez, J.M.: Improved dynamic lexicographic ordering for multi-objective optimisation. Lect. Notes Comput. Sci., \textbf{6239}, 31--40 (2010) 
-\bibitem{c44} Audet, C., Savard, G., Zghal, W.: Multiobjective optimization through a series of single-objective formulations. SIAM J. Optim. \textbf{19}(1), 188--210 (2008)
-\bibitem{c45} Sergeyev, Y.D., Kvasov, D.E.: Deterministic Global Optimization: An Introduction to the Diagonal Approach (Springer Briefs in Optimization). Springer. (2017)
-\bibitem{c46} Carr, C.R., Howe, C.W.: Quantitative Decision Procedures in Management and Economic: Deterministic Theory and Applications. New York: McGraw–Hill (1964)
-\bibitem{c47} Dam, E.R., Husslage, B., Hertog, D.: One-dimensional nested maximin designs. J. Glob. Optim. \textbf{46}(2), 287--306 (2010)
-\bibitem{c48} Gergel, V., Kozinov, E.: Efficient multicriterial optimization based on intensive reuse of search information, J.
-Glob.
-Optim., \textbf{71}(1), 73--90 (2018).
-DOI:10.1007/s10898-018-0624-3.
-\bibitem{c49} Mart\'l, L., Garc\'la, J., Berlanga, A., Molina, J.M.: A stopping criterion for multi-objective optimization evolutionary algorithms. Inf. Sci. \textbf{367--368}, 700--718 (2016)
-\bibitem{c50} Dilettoso, E., Rizzo, S.A., Salerno, N.: A weakly Pareto compliant quality indicator. Math. Comput. Appl. \textbf{22}(1), 25 (2017)
-\bibitem{c51} Wu, J., Azarm, S.: Metrics for quality assessment of a multiobjective design optimization solution set. J. Mech. Des. \textbf{123}(1), 18--25 (2001)
-\bibitem{c52} Audet, C., Bigeon, J., Cartier, D., Le Digabel, S., Salomon, L.: Performance indicators in multiobjective optimization. Technical Report G-2018-90, Les cahiers du GERAD. (2018)
-\bibitem{c53} Huband, S., Hingston, P., Barone, L., While, L.: A review of multiobjective test problems and a scalable test problem toolkit. IEEE T. Evolut. Comput. \textbf{10}(5), 477--506 (2006)
-\bibitem{c54} Hansen, N., Auger, A., Finck, S., Ros, R.: Real-Parameter Black-Box Optimization Benchmarking 2009: Experimental Setup.
-INRIA Research Report RR-6829, INRIA Saclay-Ile-de-France.
-updated February 2010 (2009)
-\bibitem{c55} Cust\'odio, A.L., Madeira, J.F.A., Vaz, A.I.F., Vicente, L.N.: Direct multisearch for multiobjective optimization. SIAM J. Optim. \textbf{21}(3), 1109--1140 (2011)
-\bibitem{c56} Brockhoff, D., Tusar, T., Auger, A., Hansen, N.: Using well-understood single-objective functions in multiobjective black-box optimization test suites.
-(2019) https://arxiv.org/abs/1604.00359
 \bibitem{c57} Cococcioni, M., Cudazzo, A., Pappalardo, M., Sergeyev, Y.D.: Solving the lexicographic multi-objective mixed-integer linear programming problem using branch-and-bound and grossone methodology.
@@
 
-\bibitem{Book2013} Strongin, R.G., Gergel, V.P., Grishagin, V.A., Barkalov, K.A.: Parallel computing in global optimization problems. MSU Press, Moscow. (2013, in Russian)
+\bibitem{c38} Zarepisheh, M., Khorram, E.: On the transformation of lexicographic nonlinear multiobjective  programs to single objective programs. Math. Methods Oper. Res. \textbf{74}, 217--231 (2011)
+\bibitem{c39} Haimes, Y.Y., Lasdon, L., Wismer, D.: On a bicriterion formulation of the problem of integrated systems identification and system optimization.
+IEEE Trans.
+Syst.
+Man Cybern.
+\textbf{1}, 296--297 (1971)
+\bibitem{c40} Chankong, V., Haimes, Y.Y.: Multiobjective Decision Making: Theory and Methodology. North-Holland Series in System Science and Engineering. New York (NY): Elsevier. (1983)
+\bibitem{c41} Ehrgott, M., Ruzika, S.: Improved $\varepsilon$-constraint method for multiobjective programming. J. Optim. Theory Appl. \textbf{138}, 375--396 (2008)
 
-\bibitem{Sergeyev2003} Sergeyev, Y.D., Pugliese, P., Famularo, D.: Index information algorithm with local tuning for solving multidimensional global optimization problems with multiextremal constraints.
-Math.
-Program., Ser.
-A \textbf{96}, 489--512 (2003)
+\bibitem{c42} Rastegar, N., Khorram, E.: Relaxation of constraints in lexicographic multiobjective programming problems.
+A Journal of Mathematical Programming and Operations Research, \textbf{64}, 2111--2129 (2015)
+\bibitem{c43} Castro-Gutierrez, J., Landa-Silva, D., P\'erez, J.M.: Improved dynamic lexicographic ordering for multi-objective optimisation. Lect. Notes Comput. Sci., \textbf{6239}, 31--40 (2010)
+
+\bibitem{c31} Gergel, V.P., Kozinov, E.A.: Efficient multicriterial optimization based on intensive reuse of search information.
+J.
+Glob.
+Optim.
+\textbf{71}(1), 73--90 (2018).
+DOI: 10.1007/s10898-018-0624-3
+\bibitem{c44} Audet, C., Savard, G., Zghal, W.: Multiobjective optimization through a series of single-objective formulations. SIAM J. Optim. \textbf{19}(1), 188--210 (2008)
+
+\bibitem{c15} Jones, D.R.: A taxonomy of global optimization methods based on response surfaces. J. Glob. Optim. \textbf{21}, 345--383 (2001)
+\bibitem{c16} Voutchkov, I., Keane, A.: Multi-objective optimization using surrogates. Comput. Intell. Optim. Adapt. Learn. Optim. \textbf{7}, 155--175 (2010)
+\bibitem{c17} Strongin, R.G.: Numerical Methods in Multiextremal Problems: Information-statistical Algorithms. Nauka, Moscow. (1978, in Russian)
+
+\bibitem{c18} Strongin, R.G., Sergeyev, Y.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht. (2nd ed. 2013, 3rd ed. 2014).
+\bibitem{c19} T\"orn, A., {\v Z}ilinskas, A.: Global Optimization. Lecture Notes in Computer Science, 350. Springer-Verlag, Berlin. (1989)
+\bibitem{c20} Horst, R., Tuy, H.: Global Optimization: Deterministic Approaches. Springer-Verlag, Berlin. (1990)
+\bibitem{c21} Zhigljavsky, A.A.: Theory of Global Random Search. Kluwer Academic Publishers, Dordrecht. (1991)
+\bibitem{c22} Pint\'er, J.D.: Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications). Kluwer Academic Publishers, Dortrecht. (1996)
+\bibitem{c23} Sergeyev, Y.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting Space-filling Curves. Springer. (2013)
+\bibitem{c24} Locatelli, M., Schoen, F.: Global Optimization: Theory, Algorithms, and Applications. SIAM. (2013)
+\bibitem{c25} Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. Princeton University Press. (2016)
+
+\bibitem{c45} Sergeyev, Y.D., Kvasov, D.E.: Deterministic Global Optimization: An Introduction to the Diagonal Approach (Springer Briefs in Optimization). Springer. (2017)
+\bibitem{c46} Carr, C.R., Howe, C.W.: Quantitative Decision Procedures in Management and Economic: Deterministic Theory and Applications. New York: McGraw–Hill (1964)
+\bibitem{c47} Dam, E.R., Husslage, B., Hertog, D.: One-dimensional nested maximin designs. J. Glob. Optim. \textbf{46}(2), 287--306 (2010)
+\bibitem{Grishagin2019} Gergel V.P., Grishagin V.A., Israfilov R.A.
+Adaptive dimensionality reduction in multiobjective optimization with multiextremal criteria.
+Lecture Notes in Computer Science, \textbf{11331}, 129--140 (2019)
 
@@
 
-\bibitem{Grishagin2019} Gergel V.P., Grishagin V.A., Israfilov R.A.
-Adaptive dimensionality reduction in multiobjective optimization with multiextremal criteria.
-Lecture Notes in Computer Science, \textbf{11331}, 129--140 (2019)
+\bibitem{c26} Lera, D., Sergeyev, Y.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and Holder constants.
+Commun.
+Nonlinear Sci.
+\textbf{23}(1-3), 328--342 (2015)
+\bibitem{c27} Gergel, V.: A unified approach to use of coprocessors of various types for solving global optimization problems. 2nd Int. Conf. on MACSI. 13--18 (2015). DOI: 10.1109/MCSI.2015.18
+\bibitem{c28} Arora, R.K.: Optimization: Algorithms and Applications. CRC Press. (2015)
+\bibitem{c29} Bazaraa, M.S., Sherali, H.D., Shetty, C.M.: Nonlinear Programming: Theory and Algorithms. John Wiley and Sons. (2006)
+
+\bibitem{Book2013} Strongin, R.G., Gergel, V.P., Grishagin, V.A., Barkalov, K.A.: Parallel computing in global optimization problems. MSU Press, Moscow. (2013, in Russian)
+\bibitem{Sergeyev2003} Sergeyev, Y.D., Pugliese, P., Famularo, D.: Index information algorithm with local tuning for solving multidimensional global optimization problems with multiextremal constraints.
+Math.
+Program., Ser.
+A \textbf{96}, 489--512 (2003)
+
+\bibitem{c30} Gergel, V.P., Kozinov, E.A.: Accelerating multicriterial optimization by the intensive exploitation of accumulated search data.
+In: 	AIP Conf.
+Proc., \textbf{1776}, 090003 (2016).
+DOI: 10.1063/1.4965367
+\bibitem{c32} Sysoyev, A., Barkalov, K., Gergel, V.: Globalizer: A novel supercomputer software system for solving time-consuming global optimization problems.
+Numer.
+Algebra, Control.
+Optim.
+\textbf{8}(1), 47--62 (2018)
+
+\bibitem{c49} Mart\'l, L., Garc\'la, J., Berlanga, A., Molina, J.M.: A stopping criterion for multi-objective optimization evolutionary algorithms. Inf. Sci. \textbf{367--368}, 700--718 (2016)
+\bibitem{c50} Dilettoso, E., Rizzo, S.A., Salerno, N.: A weakly Pareto compliant quality indicator. Math. Comput. Appl. \textbf{22}(1), 25 (2017)
+\bibitem{c51} Wu, J., Azarm, S.: Metrics for quality assessment of a multiobjective design optimization solution set. J. Mech. Des. \textbf{123}(1), 18--25 (2001)
+\bibitem{c52} Audet, C., Bigeon, J., Cartier, D., Le Digabel, S., Salomon, L.: Performance indicators in multiobjective optimization. Technical Report G-2018-90, Les cahiers du GERAD. (2018)
+\bibitem{c35} Evtushenko, Yu.G., Posypkin, M.A.: A deterministic algorithm for global multi-objective optimization. Optim. Methods Softw., \textbf{29}(5), 1005--1019 (2014)
+\bibitem{c36} {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Adaptation of a one-step worst-case optimal univariate algorithm of bi-objective Lipschitz optimization to multidimensional problems.
+Commun.
+Nonlinear Sci.
+Numer.
+Simul., \textbf{21}(1-3), 89--98 (2015)
+
+\bibitem{c53} Huband, S., Hingston, P., Barone, L., While, L.: A review of multiobjective test problems and a scalable test problem toolkit. IEEE T. Evolut. Comput. \textbf{10}(5), 477--506 (2006)
+\bibitem{c54} Hansen, N., Auger, A., Finck, S., Ros, R.: Real-Parameter Black-Box Optimization Benchmarking 2009: Experimental Setup.
+INRIA Research Report RR-6829, INRIA Saclay-Ile-de-France.
+updated February 2010 (2009)
+\bibitem{c55} Cust\'odio, A.L., Madeira, J.F.A., Vaz, A.I.F., Vicente, L.N.: Direct multisearch for multiobjective optimization. SIAM J. Optim. \textbf{21}(3), 1109--1140 (2011)
+\bibitem{c56} Brockhoff, D., Tusar, T., Auger, A., Hansen, N.: Using well-understood single-objective functions in multiobjective black-box optimization test suites.
+(2019) https://arxiv.org/abs/1604.00359
+
+\bibitem{c37} Gaviano, M., Kvasov, D.E, Lera, D., Sergeyev, Y.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
+ACM Trans.
+Math.
+Softw.
+\textbf{29}(4), 469--480 (2003)
+
+\bibitem{c33} Kvasov, D.E., Sergeyev, Y.D.: Deterministic approaches for solving practical black-box global optimization problems. Adv. Eng. Softw. \textbf{80}, 58--66 (2015)
+\bibitem{c34} Modorskii, V., Gaynutdinova, D., Gergel, V., Barkalov, K.: Optimization in design of scientfic products for purposes of cavitation problems. AIP Conf. Proc. \textbf{1738}, 400013 (2016)
 

### 2020-12-02 2021/ICCS_BarkalovUsova/iccs.tex
Initial commit for ICCS paper

--- 2019-03-19 2019/WCGO/WCGO.tex
@@
 
-%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
+%%%% ДЛЯ РУССКОГО ТЕКСТА
 %\usepackage[utf8x]{inputenc}
@@
 %
-\title{Adaptive Global Optimization Based on Nested Dimensionality 
-Reduction\thanks{This study was supported by the Russian Science Foundation, 
-project No.\,16-11-10150.}}
+\title{An algorithm for searching the global minimum of univariate discontinuous functions\thanks{This work was supported by project no. 0729-2020-0055.}}
 %
-\titlerunning{Adaptive Global Optimization}
+\titlerunning{An algorithm for searching the global minimum}
 % If the paper title is too long for the running head, you can set 
@@
 \author{Konstantin Barkalov %\orcidID{0000-0001-5273-2471}  
-\and Ilya Lebedev %\orcidID{0000-0002-8736-0652}
+\and Marina Usova %\orcidID{0000-0002-8736-0652}
 }
 %
-\authorrunning{K. Barkalov, I. Lebedev}
+\authorrunning{K. Barkalov, M. Usova}
 % First names are abbreviated in the running head.
@@
 
-In the present paper, the multidimensional multiextremal optimization 
-problems and the numerical methods for solving these ones are considered. A 
-general assumption only is made on the objective function that this one 
-satisfies the Lipschitz condition with the Lipschitz constant not known 
-a priori. The problems of this type are frequent in the applications. 
-Two approaches to the dimensionality reduction for the multidimensional 
-optimization problems were considered. The first one uses the Peano-type 
-space-filling curves mapping a one-dimensional interval onto a multidimensional 
-domain. The second one is based on the nested optimization scheme, which 
-reduces a multi-dimensional problem to a family of the one-dimensional 
-subproblems. A generalized scheme combining these two approaches has been 
-proposed. In this novel scheme, solving a multidimensional problem is 
-reduced to solving a family of problems of lower dimensionality, in 
-which the space-filling curves are used. An adaptive algorithm, in which all 
-arising subproblems are solved simultaneously has been implemented. The 
-numerical experiments on several hundred test problems have been carried out 
-confirming the efficiency of the proposed generalized scheme.
-
 \keywords{Global optimization \and Multiextremal functions \and 
@@
 \section{Introduction}
-This paper considers ``black-box'' global optimization problems of the 
-following form:
-\begin{eqnarray}\label{main_problem}
-& \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
-& D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
-\end{eqnarray}
-The objective function is assumed to satisfy the Lipschitz condition 
-\[
-\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in
- D,\; 0<L<\infty,
-\]
-with the constant $L$ unknown a priori.
-
-The multistart scheme is a well known method for solving the multiextremal problems.
-In such schemes, a grid is seeded in the search domain, the points of which are used as the starting ones for the search of 
-the extrema by some local method, and then the lowest of the found extrema is chosen. 
-The choice of the starting points performed, as a rule, on the basis of the Monte Carlo method is a special problem within 
-this approach \cite{Zhigljavsky2008}. 
-The approach is well applicable for the problems with a small number of local minima having a vide field of application, 
-but for the problems with essential multiextremality its efficiency falls drastically.
-
-At present, the genetic algorithms one way or another based on the random search concept are used widely for solving the 
-global optimization problems (see, for example, \cite{Yang2013}). 
-Because of simplicity of implementation and usage these ones have gained a large popularity. 
-However the quality of work (the number of problems from some set solved correctly can serve as a quantitative measure 
-of which) is essentially less as compared to the deterministic algorithms \cite{Kvasov2018,Sergeyev2018}.
-
-If one speaks about the deterministic global optimization methods, many ones of this class are based on various methods 
-of division of the search domain into a system of subdomains followed by the selection of the most promising subdomain 
-for placing the next trial (computing the objective function). The results in this direction are presented in 
-\cite{Evtushenko2013,Jones2009,Paulavicius2016,Zilinskas2010,Sergeyev2015}.
-
-Finally, the approach related to the reduction of the multidimensional problems to the equivalent one-dimensional ones (or 
-to a system of the one-dimensional subproblems) with subsequent solving the one-dimensional problems by the efficient 
-univariate optimization algorithms is widely used for the development of the multidimensional 
-optimization methods. Two such schemes are used: the reduction based on the Peano-type space-filling curves (evolvents) 
-\cite{Sergeyev2013,Strongin2000}, and the nested optimization scheme \cite{Grishagin2001,Strongin2000}.
-
-An adaptive reduction scheme generalizing the classical nested optimization one has been proposed in 
-\cite{Grishagin2016}. The adaptive scheme essentially enhances the optimization efficiency as compared to the base 
-prototype \cite{Grishagin2016_1}. In the present work, a generalization of the adaptive dimensionality reduction scheme 
-combining the use of the nested optimization and the Peano curves has been proposed. In this approach the nested 
-subproblems in the adaptive scheme can be the one-dimensional as well as the multidimensional ones. In the latter case, 
-the evolvents are used for the reduction of the dimensionality of the nested subproblems.
 
@@
 
-As a core problem we consider a one-dimensional multiextremal optimization 
-problem
-\begin{equation}\label{uni_problem}
-\varphi^\ast = \varphi(x^\ast)=\min{\left\{\varphi(x):x\in \left[0,1\right] 
-\right\}}
-\end{equation}
-with objective function satisfying the Lipschitz condition.
-
-Let us give the description of the global search algorithm (GSA) applied for 
-solving the above problem (according to \cite{Strongin2000}). 
-GSA involves constructing a sequence of points $x^i$, where the values of the 
-objective function $z^i=\varphi(x^i)$ are calculated. Let us call the 
-function value calculation process the \textit{trial}. 
-According to the algorithm, the first two trials are executed at the ends of 
-the interval  $[0,1]$, i.e. $x^0=0,\;x^1=1$. The function values $z^0=\varphi
-(x^0),\;z^1=\varphi(x^1)$  are computed and the number $k$ is set to 1. In 
-order to select the point of a new trial $x^{k+1}, k\geq 1,$  it is necessary 
-to perform the following steps.
-
-\textbf{Step 1.} Renumber by subscripts (beginning from zero) the points $x^i,
-\:0\leq i\leq k$, of the previous trials in increasing order, i.e.,
-\[
-0=x_0<x_1<\ldots <x_{k}=1.
-\] 
-Juxtapose to the points $x_i, 0\leq i\leq k$,  the function values $z_i=
-\varphi(x_i), 0\leq i\leq k$.
-
-\textbf{Step 2.} Compute the maximum absolute value of the first divided 
-differences 
-\begin{equation}\label{mu}
-\mu=\max_{1\leq i\leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i}
-\end{equation}
-where $\Delta_i = x_i-x_{i-1}$. If the above formula yields a zero value, 
-assume that $\mu = 1$.
-
-\textbf{Step 3.} For each interval $(x_{i-1},x_i),1\leq i\leq k$,  calculate 
-the characteristic
-\begin{equation}\label{R}
-R(i)=r\mu\Delta_i+\frac{(z_i-z_{i-1})^2}{r\mu\Delta_i}-2(z_i+z_{i-1}),
-\end{equation} 
-where $r>1$ is a predefined parameter of the method.
-
-\textbf{Step 4.} Find the interval $(x_{t-1},x_t)$ with the maximum 
-characteristic
-\begin{equation}\label{MaxR}
-R(t)=\max_{1\leq i\leq {k}}R(i).
-\end{equation}
-
-\textbf{Step 5.} Execute the new trial at the point 
-\begin{equation}\label{xk1}
-x^{k+1}=\frac{1}{2}(x_{t-1}+x_t) - \frac{z_t-z_{t-1}}{2r\mu}.
-\end{equation}
-
-The algorithm terminates if the condition $\Delta_t<\epsilon$ is satisfied; 
-here $t$ is from (\ref{MaxR}) and $\epsilon>0$ is the preset accuracy. For 
-estimation of the global solution values
-\[
-z_k^\ast=\min_{0\leq i \leq k}\varphi(x^i), \ x_k^\ast=\arg \min_{0\leq i \leq
- k}\varphi(x^i).
-\]
-are selected. 
-The theory of algorithm convergence is presented in \cite{Strongin2000}.
-
-\section{Dimensionality reduction}
-\subsection{Dimensionality reduction using Peano-type space-filling curves}
-\label{SectionPeano}
-
-The use of Peano curve $y(x)$ 
-\begin{equation}
-\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq N\right\}=\left\{
-y(x):0\leq x \leq 1 \right\}
-\end{equation}
-unambiguously mapping the interval of real axis $[0,1]$ onto an 
-$N$-dimensional cube is the first of the dimension reduction methods considered.
-Problems of numerical construction of Peano-type space-filling curves and the 
-corresponding theory are considered in detail in 
-\cite{Sergeyev2013,Strongin2000}. Here we will note that a numerically 
-constructed curve (\textit{evolvent}) is $2^{-m}$ accurate approximation of 
-the theoretical Peano curve, where $m$ is an evolvent construction parameter.
-
-By using this kind of mapping it is possible to reduce the multidimensional 
-problem~(\ref{main_problem}) to a univariate problem
-\begin{equation}
-\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]
-\right\}}.
-\end{equation}
-An important property of such mapping is preservation of boundedness of 
-function relative differences (see \cite{Sergeyev2013,Strongin2000}). If the 
-function $\varphi(y)$ in the domain $D$ satisfies the Lipschitz condition, 
-then the function $\varphi(y(x))$ on the interval $[0,1]$ will satisfy a 
-uniform H{\"o}lder condition
-\begin{equation}\label{Holder}
-\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N},
-\end{equation}
-where the H{\"o}lder constant $H$ is linked to the Lipschitz constant $L$ by 
-the relation
-\begin{equation}
-H=2L\sqrt{N+3}.
-\end{equation}
-Condition (\ref{Holder}) allows adopting the algorithm for solving the 
-one-dimensional problems presented in Section \ref{SectionCore} for solving the 
-multidimensional problems reduced to the one-dimensional ones. For this, the 
-lengths of intervals $\Delta_i$  involved into rules (\ref{mu}),(\ref{R}) of the
-algorithm are substituted by the lengths 
-\begin{equation}
-\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}
-\end{equation}
-and the following expression is introduced instead of formula (\ref{xk1}):
-\begin{equation}
-x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}
-\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N.
-\end{equation}
-
-\subsection{Nested optimization scheme}
-
-The nested optimization scheme of dimensionality reduction is based on the 
-well-known relation (see, e.g., \cite{Carr1964})
-\begin{equation}\label{nested}
-\min_{y \in D}\varphi(y) = \min_{a_1\leq y_1 \leq b_1}\min_{a_2\leq y_2 \leq 
-b_2}...\min_{a_N\leq y_N \leq b_N}\varphi(y),
-\end{equation}
-which allows replacing the solving of multidimensional problem 
-(\ref{main_problem}) by solving a family of one-dimensional subproblems 
-related to each other recursively.
-
-In order to describe the scheme let us introduce a set of reduced functions 
-as follows:
-\begin{equation}\label{nested_N}
-\varphi^N(y_1,...,y_N) = \varphi(y_1,...,y_N),
-\end{equation}
-\begin{equation}\label{nested_i}
-\varphi^i(y_1,...,y_i) = \min_{a_{i+1}\leq y_{i+1}\leq b_{i+1}} \varphi^{i+1}(
-y_1,...,y_i,y_{i+1}), 1\leq i\leq N-1.
-\end{equation}
-
-Then, according to relation (\ref{nested}), solving of multidimensional 
-problem (\ref{main_problem}) is reduced to solving a one-dimensional problem 
-\begin{equation}\label{nested_1}
-\varphi^* = \min_{a_1\leq y_1\leq b_1}\varphi^1(y_1).
-\end{equation}
-But in order to evaluate the function $\varphi^1$ at a fixed point $y_1$ it 
-is necessary to solve the one-dimensional problem of the second level
-\begin{equation}
-\varphi^1(y_1) = \min_{a_2\leq y_2\leq b_2}\varphi^2(y_1,y_2),
-\end{equation}
-and so on up to the univariate minimization of the function $\varphi^N(y_1
-,...,y_N)$ with fixed coordinates $y_1,...,y_{N-1}$ at the $N$-th level of 
-recursion.
-
-For the nested scheme presented above, a generalization (\textit{block nested 
-optimization scheme}), which combines the use of evolvents and the nested 
-scheme has been elaborated in \cite{Barkalov2016}.
-
-Let us consider vector $y$ as a vector of block variables
-\begin{equation}
-y=(y_1,y_2,...,y_N)=(u_1,u_2,...,u_M),
-\end{equation}
-where the $i$-th block variable $u_i$ is a vector of vector $y$ components, 
-taken serially, i.e., $u_1=(y_1,y_2,...,y_{N_1})$, $u_2=(y_{N_1+1},y_{N_1+2}
-,...,y_{N_1+N_2})$,..., $u_M=(y_{N-N_M+1},y_{N-N_M+2},...,y_{N})$, where $N_1+
-N_2+...+N_M=N$.
-
-Using the new variables, main relation of the nested scheme (\ref{nested}) 
-can be rewritten in the form 
-\begin{equation}\label{block_nested}
-\min_{y \in D}\varphi(y) = \min_{u_1\in D_1}\min_{u_2\in D_2}...\min_{u_M \in 
-D_M}\varphi(y),
-\end{equation}
-where the subdomains $D_i, 1 \leq i \leq M$, are projections of initial 
-search domain $D$ onto the subspaces corresponding to the variables $u_i, 1 
-\leq i \leq M$.
-
-The formulae defining the method for solving the problem (\ref{main_problem}) 
-based on relation (\ref{block_nested}), in general, are the same to the ones 
-of nested scheme (\ref{nested_N})--(\ref{nested_1}). It is only necessary to 
-substitute the original variables $y_i, 1\leq i \leq N$,  by the block 
-variables $u_i, 1\leq i \leq M$.
-
-At that, the nested subproblems 
-\begin{equation}\label{block_nested_i}
-\varphi^i(u_1,...,u_i) = \min_{u^{i+1}\in D_{i+1}} \varphi_{i+1}(u_1,...,u_i,u
-_{i+1}), 1\leq i\leq M-1,
-\end{equation}
-in the block scheme are the multidimensional ones. The dimension reduction 
-method based on Peano curves can be applied for solving these ones. It is a 
-principal difference from the initial nested scheme.
-
-\subsection{Block adaptive optimization scheme}
-
-The solving of the arising set of subproblems (\ref{nested_i}) (for the nested optimization scheme) or 
-(\ref{block_nested_i}) (for the block nested optimization scheme) can be organized in various ways. 
-A straightforward way (developed in details for the nested optimization scheme \cite{Grishagin2015,Grishagin2001} 
-and for the block nested optimization scheme \cite{Barkalov2016,Barkalov2014}) is based on solving the 
-subproblems according to the generation order. However, here a loss of a considerable part of the information 
-on the objective function takes place when solving the multidimensional problem. Another approach is the adaptive 
-scheme, in which all subproblems are solved simultaneously, that allows more complete accounting for the information on 
-the multidimensional and accelerating the process of its solving.
-
-For the case of the one-dimensional subproblems the adaptive scheme was theoretically substantiated and tested in 
-\cite{Grishagin2016,Grishagin2016_1,Grishagin2018}. The present work proposes a generalization of the adaptive 
-scheme for the case of the multidimensional subproblems. Let us give a brief description of its basic elements.
-
-Let us assume the nested subproblems (\ref{block_nested_i}) to be solved with the use of a multidimensional 
-global search algorithm described in Section \ref{SectionPeano}. Then, each subproblem (\ref{block_nested_i}) can be 
-associated with a numerical value called the characteristic of this problem. The value $R(t)$ from (\ref{MaxR}) (i.e., the 
-maximum characteristic of the intervals formed within the subproblem) can be taken as such characteristic. According to the 
-rule of computing the characteristics (\ref{R}), the higher the value of the characteristic, the more promising the 
-subproblem for continuing the search of the global minimum of the initial problem (\ref{main_problem}). Therefore, the 
-subproblem with the highest characteristic is selected for executing the next trial at each iteration. This trial either 
-computes the objective function value $\varphi(y)$ (if the selected subproblem belongs to the level $j=M$) or generates 
-new subproblems according to (\ref{block_nested_i}) when $j\leq M-1$. In the latter case, new generated problems are 
-added to current problem set, their characteristics are computed, and the process is repeated. The  optimization process is 
-finished when the  stop condition is fulfilled for the algorithm solving the root problem.
-
 \section{Results of numerical experiments}
 
-One of the well-known approaches to investigating and comparing multiextremal 
-optimization algorithms is based on the application of these methods for 
-solving a set of test problems generated randomly.
+Example of references 
 The comparison of the algorithms has been carried out using the Grishagin test problems $F_{gr}$ (see, for example, \cite{Grishagin1994}, test function 4) and the GKLS generator \cite{Gaviano2003}.
-
 In \cite{Barkalov2015,Grishagin2018} the global search algorithm (GSA) with the use of the evolvents as well as in combination 
@@
 
-In order to compare the efficiencies of the algorithms, the two criteria were used: the average number of trials and the 
-operating characteristic.
-The operating  characteristic of an algorithm is the function $P(k)$ defined as the fraction of the problems from the 
-considered series, for solving of which not more than $k$ trials have been required.
-The problem was considered to be solved, if the algorithm generated a trial point $y^k$ in the vicinity of 
-the global minimizer $y^\ast$, i.e. $\left|y^k-y^\ast\right| < \delta \left\|b-a\right\|$, where $\delta = 0.01$, $a$ and $b$ are the 
-boundaries of the search domain $D$.
-
-The first series of experiments has been carried out on the two-dimensional problems from the classes $F_{gr}$, GKLS 
-\textit{Simple}, and GKLS \textit{Hard} (100 functions from each class). Table~\ref{tab1} presents the averaged 
-numbers of trials executed by GSA with the use of evolvents ($K_e$), nested optimization scheme ($K_n$), and adaptive 
-nested optimization scheme ($K_{an}$). Fig.~\ref{fig1} and Fig.~\ref{fig2}(a,b) present the operating characteristics of the 
-algorithms obtained on the problem classes $F_{gr}$, GKLS \textit{Simple}, and GKLS \textit{Hard} respectively. The solid 
-line corresponds to the algorithm using the evolvents (GSA-E), short dashed line -- to the adaptive nested optimization scheme (GSA-AN), and 
-the long dashed line -- to the nested optimization scheme (GSA-N). The results of experiments demonstrate that GSA with the use of 
-the adaptive nested optimization scheme shows almost the same speed as compared to GSA with the evolvents, and 
-both algorithms considerably exceed the algorithm using the nested optimization scheme. Therefore further experiments 
-were limited to the comparison of different variants of the adaptive  dimensionality reduction scheme.
-
-\begin{table}
-\centering
-\caption{Average number of trials for 2D problems.}\label{tab1}
-\begin{tabular}{lccc}
-\hline\noalign{\smallskip}
- &  $\;\;\;F_{gr}\;\;\;$  &  GKLS \textit{Simple} &  GKLS \textit{Hard} \\
-\noalign{\smallskip}\hline\noalign{\smallskip}
- $K_e$ & 180  & 252 & 674 \\
- $K_n$ & 341  & 697 & 1252 \\
- $K_{an}$ & 215  & 279 & 815 \\
-\noalign{\smallskip}\hline
-\end{tabular}
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.50\textwidth]{2D.pdf}
-\caption{Operating characteristics using $F_{gr}$ class} 
-\label{fig1}
-\end{figure}
-
-\begin{figure}
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{2DSimple.pdf} \\ (a)}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{2DHard.pdf} \\ (b)}
-\end{minipage}
-\caption{Operating characteristics using 2d GKLS \textit{Simple} (a) and \textit{Hard} (b) classes}
-\label{fig2}
-\end{figure}
-
-The second series of experiments has been carried out on the four-dimensional problems from the classes GKLS 
-\textit{Simple} and GKLS \textit{Hard} (100 functions of each class). Table~\ref{tab2}  presents the averaged numbers 
-of trials executed by GSA with the use of the adaptive nested optimization scheme ($K_{an}$) and block adaptive nested 
-optimization scheme ($K_{ban}$) with two levels of subproblems of equal dimensionality $N_1=N_2=2$. 
-Note that when solving the problem of the dimensionality $N=4$ using the initial variant of the adaptive scheme, four 
-levels of one-dimensional subproblems are formed that complicates the processing of these ones.
-
-Fig.~\ref{fig3}(a,b) presents the operating characteristics of the algorithms obtained on the classes GKLS \textit{Simple} 
-and GKLS \textit{Hard} respectively. The dashed line corresponds to GSA using the adaptive nested optimization 
-scheme (GSA-AN), the solid line -- the block adaptive nested optimization scheme (GSA-BAN). 
-The results of experiments demonstrate the use of the block adaptive nested optimization scheme provides a considerable 
-gain in the number of trials (up to 35\%) as compared to the initial adaptive nested optimization scheme.
-
-\begin{table}
-\centering
-\caption{Average number of trials for 4D problems.}\label{tab2}
-\begin{tabular}{lcc}
-\hline\noalign{\smallskip}
- &    GKLS \textit{Simple} &  GKLS \textit{Hard} \\
-\noalign{\smallskip}\hline\noalign{\smallskip}
- $K_{an}$  &  21747 & 35633 \\
- $K_{ban}$ &  13894 & 31620 \\
-\noalign{\smallskip}\hline
-\end{tabular}
-\end{table}
-
-\begin{figure}
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{4DSimple.pdf} \\ (a)}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{4DHard.pdf} \\ (b)}
-\end{minipage}
-\caption{Operating characteristics using 4d GKLS \textit{Simple} (a) and \textit{Hard} (b) classes}
-\label{fig3}
-\end{figure}
-
 \section{Conclusion}
-
-In the present work, the generalized adaptive dimensionality reduction scheme for the global optimization problems 
-combining the use of Peano space-filling curves and the nested (recursive) optimization scheme has been proposed.
-For solving the reduced subproblems of less dimensionality, the global search algorithm was applied. 
-The computational scheme of the algorithm was given, main issues related to the use of the adaptive dimensionality 
-reduction scheme were considered.
-The numerical experiments have been carried out using the series of test problems in order to compare the efficiencies of 
-different dimensionality reduction schemes.  
-The result of experiments demonstrated that the use of the block adaptive nested optimization scheme can essentially reduce the 
-number of trials required to solve the problem with given accuracy. 
-Further works on the development of the global optimization methods based on this method of the 
-dimensionality reduction may be related to the use of the local estimates of the Lipschitz constant (considered, for example, in \cite{Grishagin2015,Sergeyev2016}) in different subproblems.
 

### 2020-12-10 2021/PСT_GergelKozinov/2021_PCT.tex
PCT 2021/PСT_GergelKozinov 1. Оформлен список литературы

\documentclass{svproc}

\usepackage{url}
\usepackage{hyperref}

\usepackage{adjustbox}
\usepackage{amsmath}

\begin{document}

\title{Parallel Computations for Solving Multicriteria Mixed-Integer Optimization Problems}

\author{Victor Gergel \and Evgeniy Kozinov}

\institute{Institute of Informational Technologies, Mathematics and Mechanics, Mathematical Center, Lobachevsky State University of Nizhny Novgorod, Nizhni Novgorod, Russia \\
\email{gergel@unn.ru, evgeny.kozinov@itmm.unn.ru}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
abstract.
\keywords{Global optimization $\cdot$ dimensionality reduction $\cdot$ search information $\cdot$ parallel methods of global search $\cdot$ computational complexity}
\end{abstract}

\section{Introduction}\label{sec:1}

\cite{c1,c2,c3,c4,c5,c6,c7,c8,c9,c10}
\cite{c11,c12,c13,c14,c15,c16,c17,c18,c19,c20}
\cite{c21,c22,c23,c24,c25,c26,c27,c28,c29,c30}
\cite{c31,c32,c33,c34,c35,c36,c37,c38,c39,c40}
\cite{c41,c42,c43,c44,c45}

\section*{Acknowledgements} 
This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''

\bibliographystyle{spmpsci}
%\bibliographystyle{unsrt}
\bibliography{mybibfile}

\end{document}

### 2020-12-12 2021/PСT_GergelKozinov/2021_PCT.tex
2021/PСT_GergelKozinov Перевел текст в tex до 5-го раздела

@@
 \usepackage{hyperref}
+%\usepackage{hypernat}
+\usepackage{cite}
+%\usepackage[numbers,compress]{natbib}
 
@@
 \usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{color,soul}
 
@@
 \begin{abstract}
-abstract.
-\keywords{Global optimization $\cdot$ dimensionality reduction $\cdot$ search information $\cdot$ parallel methods of global search $\cdot$ computational complexity}
+The paper discusses a new approach to solving computationally time-consuming multicriteria optimization problems, in which some variable parameters can only take on discrete values.
+Under the proposed approach, the solution of mixed-integer optimization problems is reduced to solving a family of optimization problems where only continuous parameters are used.
+All problems of the family are solved simultaneously in the time division mode, where the optimization problem for the next global search iteration is selected adaptively taking into account the search information obtained in the course of the calculations.
+The algorithms proposed enable parallel computing on high-performance computing systems.
+Computational experiments confirm that the proposed approach makes it possible to significantly reduce computation volume and time for solving complex multicriteria mixed-integer optimization problems.
+\keywords{Multicriteria optimization $\cdot$ multi-integer optimization problems $\cdot$ methods of criteria scalarization $\cdot$ global optimization $\cdot$ search information $\cdot$ parallel computing $\cdot$ computational experiment}
 \end{abstract}
@@
 \section{Introduction}\label{sec:1}
+
+Multicriteria optimization (MCO) problems arise almost every time when the best options need to be chosen when developing complex technical facilities and systems.
+The scale of demand for MCO problems determines the high intensity of research in this area -- see, for example, monographs \cite{c1,c2,c3,c4,c5} and reviews of scientific and practical results in this area \cite{c6,c7,c8,c9}.
+
+Usually, the solution of a MCO problem is a set of efficient (non-dominated) decisions for which an improvement in values with respect to some criteria cannot be achieved without deteriorating the efficiency with respect to some other criteria.
+The definition of the whole set of efficient decisions (the Pareto set), on the one hand, may require a large amount of computation and, on the other hand, may be redundant as the analysis of a large number of efficient decisions may require a significant effort for the person who makes the decision (decision maker, DM).
+Thus, it may be practically justified to find only a relatively small set of efficient decisions that can be formed according to the optimality requirements defined by the DM.
+
+By limiting the set of efficient decisions to be computed, one can achieve a noticeable reduction in the amount of calculations required.
+However, efficiency criteria may be complex, \textit{multi-extremal}, and calculating the values of these criteria can prove \textit{computationally demanding}.
+In addition, some variables may only take on discrete values.
+In such cases, MCO problems involve \textbf{significant computational complexity}, which can only be overcome by using high-performance supercomputer systems.
+
+Many different approaches have been proposed for solving MCO problems (see, for example, \cite{c3,c6,c10,c11}.
+Most commonly, various methods based on reduction of the vector criterion to a particular scalar function are used \cite{c2,c12}.
+The number of works on multicriteria mixed-integer problems is however more limited -- in most cases, the issues of discrete parameter analysis are considered in relation to scalar optimization problems (see, for example, reviews \cite{c13,c14}).
+The widely used deterministic methods for solving problems of this class are usually based on the Branch-and-Bound \cite{c15} or on the Branch-and-Reduce approaches \cite{c16}.
+A number of meta-heuristic and genetic algorithms are also known and are based, in one way or another, on the random search concept \cite{c17,c18}.
+
+This paper presents the results of research that was carried out to develop highly effective parallel methods of multicriteria optimization making use of all the search information obtained in the course of calculations \cite{c19,c20,c21}.
+A new contribution to this research area consists in the development of an approach for solving MCO problems, in which some of the variable parameters can only take discrete values.
+The proposed approach reduces the solution of mixed-integer optimization problems to a family of optimization problems where only continuous parameters are used.
+All problems of the family are solved simultaneously in the time division mode, where the selection of the optimization problem for the next iteration of the global search is performed adaptively taking into account the search information obtained in the course of the calculations.
+The algorithms developed enable efficient parallel computing on high-performance computing systems.
+
+The rest of the paper is organized as follows.
+In Section \ref{sec:2}, we give the statement of multicriteria optimization problems, present a minimax scheme for scalarization of the vector efficiency criterion and introduce the concept of multistage solution of multicriteria optimization problems.
+In Section \ref{sec:3}, we present a description of the proposed approach based on the reduction of mixed-integer optimization problems to solving a family of optimization problems with the use of only continuous parameters.
+This section also describes a dimensionality reduction scheme, which allows multidimensional optimization problems to be reduced to a one-dimensional global search problem.
+Section \ref{sec:4} presents a parallel algorithm for solving multicriteria multi-integer optimization problems, which is used in the framework of the proposed approach.
+Section \ref{sec:5} contains the results of numerical experiments confirming that the proposed approach is quite promising.
+In conclusion, the results obtained are discussed and possible main directions for continuing the research are outlined.
+
+\section{Problems of Multicriteria Mixed-Integer Optimization}\label{sec:2}
+
+The problem of multicriteria mixed-integer optimization (MCOmix) can be formulated as follows
+\begin{equation}\label{eq:1}
+f(y,u)\to \min, y\in D,u \in U,
+\end{equation}
+where $f(y,u)=(f_1 (y,u),f_2 (y,u), \dots,f_s (y,u))$ is the vector criterion of efficiency in which the variable parameters belong to two different types:
+\begin{itemize}
+	\item continuous parameters $y=(y_1,y_2,\dots,y_n)$, whose domain of possible values is represented by an $N$-dimensional hyperparallelepiped:
+\begin{equation}\label{eq:2}
+D=\{ y\in R^n: a_i \leq y_i \leq b_i, 1 \leq i \leq n \}, 
+\end{equation}
+for specified vectors $a$ and $b$.
+	\item discrete parameters  $u=(u_1,u_2, \dots ,u_m)$, each of which can only take on a fixed (discrete) set of values:
+\begin{equation}\label{eq:3}
+U=U_1 \times U_2 \times \dots \times U_m=\{w_k=<w_{1k},w_{2k}, \dots, w_{mk}>: 1 \leq k \leq l\}, w_{ik} \in U_i, 
+\end{equation}
+where $U_i=\{v_{i1},v_{i2}, \dots ,v_{il_i} \}$, $1 \leq i \leq m$, is a set of $l_i>0$ admissible discrete values for the parameter $u_k$, i.e., the whole set of all possible discrete parameter values $U$ contains
+\begin{equation}\label{eq:4}
+  l=\prod\limits^m_{i = 1}{l_i}
+\end{equation}
+different elements (tuples $w_k$,$1 \leq k \leq l$).
+Without loss of generality, it will be assumed in further consideration that the criteria $f_i (y,u)$, $1 \leq i \leq s$, are non-negative and their reduction corresponds to an increase in the efficiency of the decisions selected.
+\end{itemize}
+
+In the most complex variant, the criteria $f_i(y,u)$, $1 \leq i \leq s$, may be multi-extremal, and the procedures for calculating their values can be computationally time-consuming.
+It is also assumed that the criteria $f_i (y,u)$, $1 \leq i \leq s$, meet the Lipschitz condition
+\begin{equation}\label{eq:5}
+  |f_i (y_1,u_1 )-f_i (y_2, u_2 )| \leq L_i \|(y_1, u_1)-(y_2, u_2)\|, 1 \leq i \leq s,
+\end{equation}
+where $L_i$ is the Lipschitz constant for the criterion $f_i (y,u)$, $1 \leq i \leq s$,  and $\|*\|$ denotes the Euclidean norm in $R^N$.
+
+Efficiency criteria in an MCO problem are usually controversial and the parameters $(y^*,u^*) \in D \times U$ with the best values for all criteria at the same time may not be available.
+In such situations, it is a common approach for MCO problems to find efficient (non-dominated) decisions, for which an improvement in the values of some criteria leads to a deterioration of efficiency indicators for other criteria.
+Obtaining the whole set of efficient decisions (the Pareto set) can require a lot of computation and, as a result, another approach is often used -- to find only a relatively small set of efficient decisions defined according to the decision maker's requirements.
+
+A commonly used approach to obtaining individual efficient decisions is to transform a vector criterion into some common scalar efficiency index\footnote{It should be noted that this approach ensures that a wide range of already existing global optimization methods can be used to solve MCO problems.}
+\begin{equation}\label{eq:6}
+  \min{F(\alpha,y,u)}, y \in D, u \in U,
+\end{equation}
+where $F$ is an objective function generated by scalarization of the criteria $f_i$, $1 \leq i \leq s$, $\alpha$ is the vector of the parameters of the criteria convolution applied, while $D$ and $U$ are the domains of possible parameter values from (\ref{eq:2}-\ref{eq:3}).
+By virtue of (\ref{eq:5}), the function $F(\alpha, y, u)$ also satisfies the Lipschitz condition with some constant $L$, i.e.
+\begin{equation}\label{eq:7}
+  |F(\alpha, y_1, u_1)-F(\alpha, y_2, u_2)| \leq L\|(y_1,u_1)-(y_2,u_2)\|.
+\end{equation}
+
+To construct a common scalar efficiency index $F(\alpha,y,u)$ from (\ref{eq:6}), one of the most frequently used scalarization methods is to use a minimax criteria convolution \cite{c2,c5}:
+\begin{equation}\label{eq:8}
+\begin{matrix}
+  F(\lambda,y,u)=\max{(\lambda_i f_i (y,u),1\leq i \leq s)},\\
+  \lambda=(\lambda_1,\lambda_2, \dots, \lambda_s)\in \Lambda \subset R^s: \sum_{i=1}^{s}\lambda_i=1, \lambda_i \geq 0, 1 \leq i \leq s.
+\end{matrix}
+\end{equation}
+
+It should be noted that due to the possible changes in the requirements for optimality in the process of calculations it may be necessary to change the parameters of the convolution $\lambda$ from (\ref{eq:8}).
+Such variations yield a set of scalar global optimization problems (\ref{eq:6})
+\begin{equation}\label{eq:10}
+\mathbb{F}_T=\{F(\alpha_t,y):1 \leq t \leq T\},
+\end{equation}
+which is necessary for solving the MCOmix problem.
+This set of problems can be formed sequentially in the course of calculations; the problems of the set can be solved strictly sequentially or simultaneously in time division mode.
+In addition, the problems of the set $\mathbb{F}_T$ can be solved in parallel using high-performance computing systems.
+The possibility of forming the set $\mathbb{F}_T$ defines a \textit{new approach to multistage solution of multicriteria optimization} (MMCO) problems -- see, for example, \cite{c22}.
+
+\section{The Approach: Unrolling Mixed-Integer Optimization Problems and Dimensionality Reduction}\label{sec:3}
+
+In the presence of discrete parameters, solving multicriteria optimization  problems becomes considerably  more complicated -- in many cases, it becomes necessary to calculate criterion values for all possible values of discrete parameters.
+The proposed approach to improving the efficiency of solving MCOmix problems is based on the following two basic ideas: unrolling mixed-integer optimization problems \cite{c23} and dimensionality reduction \cite{c24,c25}.
+
+\subsection{Simultaneous Solution of Mixed-Integer Optimization Problems} \label{subsec:31}
+
+To solve the global optimization problem (\ref{eq:6}), a two-stage nested optimization scheme can be used
+\begin{equation}\label{eq:11}
+\begin{matrix}
+  F(\alpha,y^*,u^*)=\min{F(\alpha,y,u)} \min_{y \in D} \min_{u \in U}{F(\alpha,y,u)} =\\ 
+  \min_{y \in D}(F(\alpha,y,w_1 ),F(\alpha,y,w_2 ), \dots, F(\alpha,y,w_l )),w_i\in U,1 \leq i \leq l.
+\end{matrix}
+\end{equation}
+
+In the computational scheme (\ref{eq:11}), for any value of continuous parameters $y \in D$, the values of the function $F(\alpha,y,u)$ are calculated for all possible values of discrete parameters $u \in U$.
+However, the smallest value of the function $F(\alpha,y,u)$ is achieved only with one specific value of discrete parameters $u^*$, so the calculation of function $F(\alpha,y,u)$ values for other values of discrete parameters is redundant.
+Thus, for example, for the problem
+\begin{equation}\label{eq:12}
+ \min\{u^2(sin(x)+sin(10x/3)):x\in[2.7,7.5],u\in \{1,2\}\}
+\end{equation}
+having one continuous and one discrete parameter, the plots of the  function $F(\alpha,y,u)$ for different values of the discrete parameter have \hl{form from }Fig.~\ref{fig:1} 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\linewidth]{fig1}
+  \caption{The function $F(\alpha,y,u)$ of the problem (\ref{eq:12}) plotted for different values of the discrete parameter (left, $u=1$,  right, $u=1$)}
+  \label{fig:1}
+\end{figure}
+and, as one can see, there is no need to calculate the value of the function $F(\alpha,y,u)$ for $u=1$.
+
+It is possible to increase the efficiency and, accordingly, to reduce computational complexity when solving  the problem (\ref{eq:6}) by excluding (or at least by reducing) the calculation of the function $F(\alpha,y,u)$ for discrete parameters $u \neq u^*$.
+In some rare situations, the information about the values of the discrete parameters that do not allow the minimum value of the function $F(\alpha,y,u)$ to be reached, may be known a priori.
+In most cases, such information can be obtained only in the process of calculations based on the obtained search information.
+
+Assume that 
+\begin{equation}\label{eq:13}
+ \omega_k=\{ y^i,u^i,z^i=F(\alpha,y^i,u^i ): 1\leq i\leq k \}
+\end{equation}
+is the search information obtained in the process of calculation after global search iterations.
+Then the procedure for adaptive estimation of the discrete parameter value, at which the minimum value  of the function $F(\alpha,y,u)$ is expected to be obtained, can be determined using the decision rule
+\begin{equation}\label{eq:14}
+	\theta =\Theta(\omega_k ),1 \leq \theta \leq l, 
+\end{equation}
+ that determines the most promising value of the discrete parameter $u \in U$ at each step of the global search.
+
+With the decision rule $\Theta$, the general computational scheme for solving the problem can be presented as follows.
+
+\textit{Rule 1.} Apply the decision rule $\Theta$ from (\ref{eq:14}) to available search information $\omega_k$ from (\ref{eq:13}),  and determine the value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.
+ .
+\textit{Rule 2.} Determine the point $y \in D$ of the global search iteration to be performed with a fixed value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.
+
+\textit{Rule 3.} Check the stopping condition for the computation.
+If the required accuracy of the global search is not achieved, it is necessary to supplement the search information $\omega_k$ from (\ref{eq:13}) with the results of the current iteration and execute rule 1.
+
+   The above computational scheme will be presented in more detail after the description of the global search algorithm.
+
+\subsection{Dimensionality Reduction of Mixed-Integer Optimization Problems} \label{subsec:32}
+
+It should be noted that the computational complexity of solving global optimization problems increases exponentially with increasing dimensionality.
+The problem of increasing computational complexity was even called the ``dimensionality curse''.
+In particular, when applying the computational scheme discussed in paragraph \ref{subsec:31}, it is necessary to accumulate and analyze multidimensional search information $\omega_k$ from (\ref{eq:13}).
+This computational complexity can be significantly reduced by reducing the optimization problems to be solved by using Peano \textit{curves} or \textit{evolvents} $y(x)$, which unmbiguously and continuously map a segment [0,1] on an $n$-dimensional hyper-cube $D$ -- see, for example, \cite{c24,c25}.
+As a result of this reduction, the multidimensional global optimization problem (\ref{eq:6}) is reduced to a one-dimensional problem:
+\begin{equation}\label{eq:15}
+ \min_{y\in D, u \in U}F(\alpha,y,u) = \min_{x \in [0,1]}(F(\alpha,y(x),w_1 ),F(\alpha,y(x),w_2 ),\dots,F(\alpha,y(x),w_l ))
+\end{equation}
+
+It should be noted that the one-dimensional functions $F(\alpha,y(x),u)$ resulting from the reduction, with a fixed value of discrete parameters $u \in U$ satisfy the uniform H{\" o}lder condition, i.e.
+\begin{equation}\label{eq:16}
+|F(\alpha, y(x_1),u_1)-F(\alpha, y(x_2),u_2)| \leq H|x_1-x_2 |^{1/n},
+\end{equation}
+where the constant $H$ is determined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:7}) and $n$ is the dimensionality of the optimization problem (\ref{eq:1}).
+
+As a result of dimensionality reduction, the search information $\omega_k$ from (\ref{eq:13}) obtained in the calculation can be presented in the form 
+\begin{equation}\label{eq:17}
+A_k=\{(x_i, u_i, z_i, f_i = f(y(x_i), u_i): 1 \leq i \leq k \},
+\end{equation}
+where $x_i$, $u_i$, $1 \leq i \leq k$,  are the points of the global search iterations, $z_i$, $f_i$, $1 \leq i \leq k$,  are the values of the scalar criterion $F(\alpha,y(x),u)$ and the criteria $f_i (y)$, $1 \leq i \leq s$, calculated at the points $x_i$, $u_i$, $1 \leq i \leq k$.
+Note that the data in the set $A_k$ are arranged in ascending order\footnote{Data ordering is reflected by using a subscript.} of the points $x_i$, $1 \leq i \leq k$, i.e.
+\begin{equation}\label{eq:18}
+ x_1< x_2< \dots < x_k
+\end{equation}
+for a more efficient execution of global search algorithms.
+
+The use of dimensionality reduction makes it possible to combine (concatenate) one-dimensional functions into a single one-dimensional function $\Phi(\alpha,y(x))$, defined over the segment $[0,l]$ (see Fig.~\ref{fig:1}).
+
+\begin{equation}\label{eq:19}
+ \Phi(\alpha,y(x)) = 
+\begin{cases}
+F(\alpha,y(x),w_1 ),x \in [0,1], \\
+F(\alpha,y(x),w_2 ),x \in [1,2], \\
+\dots \\
+F(\alpha,y(x),w_l ),x \in [l-1,l],
+\end{cases}
+\end{equation}
+where $l$ from (\ref{eq:4}) is the number of different variants of the values of discrete parameters $u \in U$.
+The mapping of the extended segment $[0,l]$ onto the domain $D$ of the continuous parameter values from (\ref{eq:2}) can be determined as follows
+\begin{equation}\label{eq:20}
+Y(x)=y(x-E(x)),x\in[0,l]
+\end{equation}
+($E(x)$ denotes the operation of taking the integral part of the number $x$).
+It should be noted that the function $\Phi(\alpha,y(x))$ is discontinuous at the points $x_i$, $1 \leq i \leq l-1$,  further on, the values of the function $\Phi(\alpha,y(x))$ at these points are considered undefined are not used in the calculations.
+
+\section{Parallel Computation for Solving Mixed-Integer Optimization Problems}\label{sec:4}
+
+In the general case, the problem of minimizing the function $F(\lambda,y,u)$ from (\ref{eq:8}) is one of  global optimization problems.
+The solution of such problems involves the construction of grids covering the search domain $D$ -- see, for example, \cite{c24,c25,c26,c27,c28,c29,c30,c31}.
+
+The proposed approach uses the algorithm of global mixed-integer search (AGMIS) to minimize the function $F(\lambda,y,u)$.
+This algorithm expands the possibilities of multi-extremal optimization methods developed within the framework of the information-statistical theory of global search \cite{c19,c20,c21,c22,c23,c32,c33,c34,c35,c36,c37,c38,c39} to minimize the reduced one-dimensional function $\Phi(\alpha,y(x))$ from (\ref{eq:19}).
+
+The AGMIS general computational scheme can be presented as follows (see also \cite{c21}).
+
+At the initial iteration of AGMIS, the value of the minimized function is calculated at some arbitrary point of the interval $(0,l)$(further, obtaining the function value will be called a \textit{trial}).
+Next, let $k$, $k > 1$, global search iterations be performed.
+Selection of the test points $(k+1)$ for the next iteration is determined by the following rules.
+
+\textit{Rule 1.} For each interval $(x_{i-1}, x_i)$,$1<i \leq k$, calculate the value $R(i)$, further referred to as the interval \textit{characteristic}.
+
+\textit{Rule 2.} Determine the interval $(x_{t-1}, x_t)$, to which the maximum characteristic
+\begin{equation}\label{eq:21}
+R(t)=\max\{R(i): 1<i\leq k\}.
+\end{equation}
+corresponds.
+
+\textit{Rule 3.} Perform a new trial at $x^{k+1}$ in the interval $(x_{t-1}, x_t)$ with the maximum characteristic
+\begin{equation}\label{eq:22}
+x^{k+1} \in (x_{t-1}, x_t).
+\end{equation}
+(the values of the discrete parameters u are given according to (\ref{eq:19})).
+
+ The stopping condition under which the trials are stopped is determined by the following condition
+\begin{equation}\label{eq:23}
+(x_t - x_{t-1})^{1/n} \leq \varepsilon, 
+\end{equation}
+where $t$ is from (\ref{eq:21}), $n$ is the dimensionality of the problem being solved from (\ref{eq:1}), and $\varepsilon > 0$ is the specified accuracy of the problem solution.
+If the stopping condition is not satisfied, the number of iteration $k$ is increased by one, and a new global search iteration is executed.
+
+   To explain the computation scheme under consideration, the following may be noted.
+The calculated characteristics $R(i)$, $1<i \leq k$, can be interpreted as some measure of the importance of intervals in terms of the presence of the global minimum point.
+It explains the interval selection scheme for the next trial -- the point of each subsequent trial $x^{k+1}$ from (\ref{eq:22}) is chosen in the range with the maximum value of the interval characteristic (i.e., in the interval  where the global minimum is most likely to be located).
+
+It should also be noted that the AGMIS computational scheme considered above refines the general scheme of global search for mixed-integer optimization problems from paragraph \ref{subsec:31}.
+Thus, the choice of the function $F(\alpha, y(x), w_i)$, $1\leq i\leq l$, for the next iteration is provided by the procedure of finding the interval with the maximum characteristic.
+
+A full description of multi-extremal optimization algorithms and conditions for their convergence developed within the framework of the information-statistical theory of global search are given in \cite{c24}.
+Thus, with a proper numerical estimation of the H{\"o}lder constants $H$ from (\ref{eq:16}), the AGMIS algorithm converges to all available points of the global minimum of the minimized function $F(\lambda, y, u)$.
+
+An example of using the AGMIS algorithm for the problem (\ref{eq:12}) within the framework of the proposed approach is shown in Fig.~\ref{fig:1}.
+In this example, the function value $F(\lambda, y, u)$ for the discrete parameter value $u=1$ was calculated only 7 times, and for the value $u=2$, 17 times, i.e.
+the global search iteration was performed mainly for the value of the discrete parameter, which allows the smallest value of the function $F(\lambda, y, u)$ to be achieved.
+
+Now, let us return to the initial problem statement (\ref{eq:1}) and remind that for solving the MCOmix problem it may be necessary to solve the set of problems $\mathbb{F}_T$ from (\ref{eq:10}).
+In solving this set of problems with the help of the AGMIS algorithm, one more key property of the proposed approach is manifested -- the results of all earlier  calculations of criteria values can be carried into the values of the next optimization problem $F(\alpha, y(x), u)$ from (\ref{eq:6}) to be solved with new values of $\alpha'$ from (\ref{eq:8}) without repeating any time-consuming calculations of criteria values, i.e.
+\begin{equation}\label{eq:24}
+ z_i'=F(\alpha',y(x_i), u), 1 \leq i \leq k.
+\end{equation}
+
+Thus, all search information $A_k$ from (\ref{eq:17}), recalculated according to (\ref{eq:24}), can be reused to continue solving the next problem $F(\alpha, y(x), u)$.
+This reuse of search information can provide a significant reduction in the amount of computations performed for each subsequent problem of the set $\mathbb{F}_T$ from (\ref{eq:10})  that may only require a relatively small number of global search iterations.
+This  has been confirmed  by our computational experiments -- see Section \ref{sec:5}.
+
+The AGMIS algorithm, supplemented with the ability to reuse search information in MCOmix problems, will be further referred to as the Algorithm for Multicriteria Mixed-Integer Search (AMMIS).
+
+The final stage in improving the efficiency of solving MCOmix problems in the proposed approach is to organize parallel computing on modern high-performance systems.
+Unfortunately, attempts to develop parallel variants of AGMIS and AMMIS algorithms using existing parallelization methods have not succeeded.
+For example, data parallelization (splitting the computation domain between available computating elements) results in only one processor processing a subdomain of the search domain  containing the sought globally optimal solution of the problem, with other processors performing redundant calculations.
+In \cite{c24,c32} a new approach to parallelizing calculations for solving global optimization problems is proposed: parallelism of calculations is provided by organizing simultaneous calculation of the values of the minimized function $F(\alpha,y,u)$ from (\ref{eq:8}) at several different points of the search domain $D$.
+This approach makes it possible to parallelize the most time-consuming part of the global search process, and, due to its general nature -- it can be applied to almost any global search method for a wide variety of global optimization problems.
+
+Applying this approach and taking into account the interpretation of the characteristics $R(i)$, $1<i \leq k$, of search intervals $(x_{i-1}, x_i)$, $1<i \leq k$, from (\ref{eq:21}) as a measure of the intervals importance in terms of containing the global minimum point, a parallel version of the AGMIS algorithm can be obtained with the following generalization of rules (\ref{eq:21})-(\ref{eq:22}) \cite{c32,c40}:
+
+\textit{Rule 2$'$}. Arrange the characteristics of intervals in descending order
+\begin{equation}\label{eq:25}
+ R(t_1) \geq R(t_2) \geq \dots \geq R(t_{k-2}) \geq R(t_{k-1})
+\end{equation}
+and select $p$ intervals with numbers $t_j$, $1 \leq j \leq p$, having maximum values of their characteristics ($p$ is the number of processors (cores) used for parallel computations).
+
+\textit{Rule 3$'$}.
+Perform new trials (calculate the values of the minimized function $F(\alpha,y(x),u)$ at the points $x^{k+j}$, $1 \leq j \leq p$ located in the intervals with the maximum characteristics from (\ref{eq:25}).
+
+The stopping condition for the algorithm (\ref{eq:23}) must be checked for all the intervals in which the next trials are performed
+\begin{equation}\label{eq:26}
+(x_{t_j} - x_{t_j-1})^{1/n} \leq \varepsilon, 1 \leq j \leq p. 
+\end{equation}
+
+As in previous cases, if the stopping condition is not fulfilled, the number of iteration $k$ is increased by $p$, and a new global search iteration is executed.
+
+This parallel variant of the AGMIS algorithm will be further referred to as Parallel Algorthm for  Global Mixed-Integer Search (PAGMIS), and the parallel variant of the AMMIS algorithm will be referred to as PAMMIS, respectively.
+
+To assess the efficiency of the parallel algorithms, a large series of computational experiments has been carried out confirming that the proposed approach allows for a significant reducion in the amount of computation and time when solving complex multicriteria mixed-integer optimization problems -- see Section \ref{sec:5}.
+
+\section{S5}\label{sec:5}
 
@@
 \section*{Acknowledgements} 
-This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''
+\hl{This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''}
 

### 2020-12-12 2021/PСT_GergelKozinov/2021_PCT.tex
2021/PСT_GergelKozinov/2021_PCT.tex Мелкие правки
2021/PСT_GergelKozinov/2021_PCT.tex
Вставил в текст статьи результаты экспериментов и заключение.

@@
 \textit{Rule 1.} Apply the decision rule $\Theta$ from (\ref{eq:14}) to available search information $\omega_k$ from (\ref{eq:13}),  and determine the value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.
- .
+
 \textit{Rule 2.} Determine the point $y \in D$ of the global search iteration to be performed with a fixed value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.
@@
 
-\textit{Rule 2$'$}. Arrange the characteristics of intervals in descending order
+\textit{Rule 2 $'$}. Arrange the characteristics of intervals in descending order
 \begin{equation}\label{eq:25}
@@
 
-\textit{Rule 3$'$}.
+\textit{Rule 3 $'$}.
 Perform new trials (calculate the values of the minimized function $F(\alpha,y(x),u)$ at the points $x^{k+j}$, $1 \leq j \leq p$ located in the intervals with the maximum characteristics from (\ref{eq:25}).
@@
 
-\section{S5}\label{sec:5}
+\section{Results of numerical experiments}\label{sec:5}
 
-\cite{c1,c2,c3,c4,c5,c6,c7,c8,c9,c10}
-\cite{c11,c12,c13,c14,c15,c16,c17,c18,c19,c20}
-\cite{c21,c22,c23,c24,c25,c26,c27,c28,c29,c30}
-\cite{c31,c32,c33,c34,c35,c36,c37,c38,c39,c40}
-\cite{c41,c42,c43,c44,c45}
+Numerical experiments were performed on the supercomputers Lobachevsky (University of Nizhni Novgorod), Lomonosov (Moscow State University), MVS-10P (Joint Supercomputer Center of RAS) and supercomputers Endeavor.
+The numerical results were obtained by using the following computational nodes: 2 Intel Xeon Platinum 8260L, 2.4 GHz, 256 GB RAM (i.e.
+a total of 48 CPU cores were available on each node).
+The executable program code was built by using the Intel Parallel Studio XE 2019 software package.
+The numerical experiments were performed using the Globalizer system \cite{c41}.
+
+The multiextremal optimization algorithms used in the framework of the proposed approach have demonstrated their effectiveness in numerical experiments and have been widely used in solving practical global search problems -- see, for example, \cite{c20,c21,c22,c23}.
+Below, we present the results of numerical experiments obtained earlier that prove the effectiveness of this approach \cite{c21}.
+In the experiments, the following bi-criteria test problem proposed in \cite{c42} was solved:
+\begin{equation}\label{eq:27}
+f_1(y)=(y_1-1) y_2^2+1,f_2 (y)=y_2, 0 \leq y_1, y_2 \leq 1.
+\end{equation}
+
+The solution of the MCO problem consisted in constructing a numerical approximation of the Pareto set.
+To assess the quality of approximation, the completeness and uniformity of the Pareto set coverage was compared using the following two indicators \cite{c42,c43}:
+\begin{itemize}
+	\item The hypervolume index (HV). This indicator evaluates the approximation of the Pareto set in terms of completeness (a higher value corresponds to a more complete coverage of the Pareto domain).
+  \item The distribution uniformity index (DU). This indicator evaluates the uniformity of the Pareto domain coverage (a lower value corresponds to a more uniform coverage of the Pareto domain).
+\end{itemize}
+
+Five multicriteria optimization algorithms were compared in this experiment: the Monte-Carlo (MC) method \cite{c42}, the genetic algorithm SEMO from the PISA library \cite{c42}, the Non-uniform coverage (NUC) method \cite{c42}, the bi-objective Lipschitz optimization (BLO) method \cite{c43}, and the AMMIS method developed by the authors within the framework of the proposed approach.
+
+For AMMIS, 50 global optimization problems with different sets of convolution coefficients were solved.
+The accuracy of the method $\varepsilon=0.05$ and the reliability parameter $r=3.0$ from (\ref{eq:23}) were used when solving the series of problems.
+The results of the experiments are presented in full in Table \ref{tab:1}.
+
+\begin{table}[ht]
+\centering
+\caption{Comparison of the efficiency of multicriteria optimization algorithms}
+\label{tab:1}
+\begin{tabular}{cccccc}
+\hline
+Solution method     & MC    & SEMO  & NUC   & BLO   & \textbf{AMMIS} \\ \hline
+Number of method iterations  & 500   & 500   & 515   & 498   & \textbf{370}   \\
+\begin{tabular}[c]{@{}c@{}}Number of points found in the Pareto domain\end{tabular} & 67    & 104   & 29    & 68    & \textbf{100}    \\
+HV index  & 0.300 & 0.312 & 0.306 & 0.308 & \textbf{0.316} \\
+DU index  & 1.277 & 1.116 & 0.210 & 0.175 & \textbf{0.101} \\ \hline
+\end{tabular}
+\end{table}
+
+As revealed in the experiments, the AMMIS algorithm has a noticeable advantage over the multicriteria optimization methods considered here even when solving relatively simple MCO problems.
+
+The results of computational experiments to solve 100 MCOmix problems of the set $\mathbb{F}_T$ from (10) are given below.
+Each problem in the set $\mathbb{F}_T$ contained two criteria.
+Each criterion was formed using the functions obtained by the GKLS generator \cite{c44}, where some parameters were discrete \cite{c45}.
+To generate MCOmix problems, 4 continuous parameters $y$ from (\ref{eq:2}) and 5 discrete parameters $u$ from (\ref{eq:3}) were used, each of them only having two values (i.e., $n=4$, $m=5$, $l=32$).
+
+To solve each MCOmix problem, 50 uniformly distributed values of convolution coefficients $\lambda \in \Lambda$ from (\ref{eq:8}) were used.
+To estimate the values of HV and DU efficiency indicators, an approximation of the Pareto domain was calculated for each MCOmix problem  using a uniform grid in the domain of variable values $D\times U$ from (\ref{eq:2})-(\ref{eq:3}).
+The PAMMIS algorithm was used for the following parameter values: accuracy $\varepsilon=0.05$ from (\ref{eq:26}) and reliability $r=5.6$ (the $r$ parameter is used in PAMMIS to construct estimates of the H{\"o}lder constant $H$ from (\ref{eq:16})).
+
+Table \ref{tab:2} gives the results of numerical experiments, averaged by the number of MCOmix problems solved.
+The first column shows the number of cores used.
+The second column contains information on the average number of trials (computation of criteria values) performed during the global search.
+The third column shows the resulting speedup (reduction in the number of trials) when using parallel computations.
+The last two columns contain the values of HV and DU indicators.
+
+\begin{table}[ht]
+\centering
+\caption{Efficiency of the PAMMIS parallel algorithm in solving MCOmix problems}
+\label{tab:2}
+\begin{tabular}{cccccc}
+\hline
+Number of computing cores    & Number of iterations    & Speedup  & DU index&   & HV index  \\\hline
+\multicolumn{6}{c}{The   Pareto domain estimate obtained by exhaustive search}        \\\hline
+48                           & 1 706 667               & -        & 20,6 & & 29 784,9  \\\hline
+\multicolumn{6}{c}{The Pareto    domain  estimate obtained using the   PAMMIS method} \\\hline
+1                            & 86 261,6                & 1,0      & 13,8 & & 30 212,3  \\
+6                            & 15 406,2                & 5,6      & 14,4 & & 30 317,6  \\
+12                           & 6 726,1                 & 12,8     & 16,2 & & 30 248,3  \\
+18                           & 5 972,7                 & 14,4     & 14,3 & & 30 551,1  \\
+24                           & 3 657,4                 & 23,6     & 13,9 & & 30 092,3  \\
+30                           & 3 162,3                 & 27,3     & 14,9 & & 30 443,9  \\
+36                           & 2 888,7                 & 29,9     & 14,6 & & 30 528,7  \\
+42                           & 2 028,8                 & 42,5     & 15,6 & & 30 046,1  \\
+48                           & 1 767,6                 & 48,8     & 15,3 & & 30 255,5 \\ \hline
+\end{tabular}
+\end{table}
+
+The results of the computational experiments show that the PAMMIS algorithm is scalable as the  speedup of parallel computations increases almost linearly with the number of computing cores used.
+The obtained values of HV and DU indicators show that the Pareto set estimate using the PAMMIS algorithm is calculated with a greater accuracy and requires a significantly smaller number of trials compared to the results obtained using uniform grids in the domain of variable values $D \times U$.
+
+\section{Conclusion}
+
+In this paper, a new approach is proposed to solving computationally intensive optimization problems, where some variable parameters can only take on discrete values (MCOmix).
+It is also assumed that the optimality criteria may be multiextremal, and the computing of criterion values may involve a large amount of calculations.
+Due to the high computational complexity of this class of problems, parallel solution methods need to be developed, relying on the efficient use of high-performance computing systems.
+
+In the framework of the proposed approach, solution of mixed-integer optimization problems is reduced to solving a family of global search optimization problems that use only continuous parameters.
+All problems of the family are solved simultaneously in the time division mode, where the optimization problem for the next global search iteration is selected adaptively taking into account the search information obtained in the course of computations.
+The results of numerical experiments prove that  this approach can significantly reduce computational intensity when solving MCOmix problems in a multistage procedure.
+
+In conclusion, it may be noted that the proposed approach is quite promising and requires continued investigations.
+First of all, it is necessary to continue numerical experiments to solve multicriteria mixed optimization problems with a greater number of efficiency criteria and having a greater dimensionality.
+The possibility of organizing parallel computations using high-performance systems with distributed memory should be also evaluated.
 

### 2020-12-13 2021/PСT_GergelKozinov/2021_PCT.tex
1
2021/PСT_GergelKozinov/2021_PCT.tex
Исправил замечания по основному тексту статьи.
Список литературы пока еще нуждается в правках.

@@
 \title{Parallel Computations for Solving Multicriteria Mixed-Integer Optimization Problems}
+\titlerunning{Parallel Computations for Solving MCO Mixed-Integer Problems}
 
@@
 Under the proposed approach, the solution of mixed-integer optimization problems is reduced to solving a family of optimization problems where only continuous parameters are used.
-All problems of the family are solved simultaneously in the time division mode, where the optimization problem for the next global search iteration is selected adaptively taking into account the search information obtained in the course of the calculations.
+All problems of the family are solved simultaneously in the time-shared mode, where the optimization problem for the next global search iteration is selected adaptively taking into account the search information obtained in the course of the calculations.
 The algorithms proposed enable parallel computing on high-performance computing systems.
 Computational experiments confirm that the proposed approach makes it possible to significantly reduce computation volume and time for solving complex multicriteria mixed-integer optimization problems.
-\keywords{Multicriteria optimization $\cdot$ multi-integer optimization problems $\cdot$ methods of criteria scalarization $\cdot$ global optimization $\cdot$ search information $\cdot$ parallel computing $\cdot$ computational experiment}
+\keywords{Multicriteria optimization $\cdot$ mixed-integer optimization problems $\cdot$ methods of criteria scalarization $\cdot$ global optimization $\cdot$ search information $\cdot$ parallel computing $\cdot$ computational experiment}
 \end{abstract}
@@
 
-Multicriteria optimization (MCO) problems arise almost every time when the best options need to be chosen when developing complex technical facilities and systems.
+Multicriteria optimization (MCO) problems arise almost every time when the optimal options need to be chosen when developing complex technical devices and systems.
 The scale of demand for MCO problems determines the high intensity of research in this area -- see, for example, monographs \cite{c1,c2,c3,c4,c5} and reviews of scientific and practical results in this area \cite{c6,c7,c8,c9}.
@@
 By limiting the set of efficient decisions to be computed, one can achieve a noticeable reduction in the amount of calculations required.
-However, efficiency criteria may be complex, \textit{multi-extremal}, and calculating the values of these criteria can prove \textit{computationally demanding}.
+However, efficiency criteria may be complex, \textit{multiextremal}, and calculating the values of these criteria can prove \textit{computationally demanding}.
 In addition, some variables may only take on discrete values.
@@
 This paper presents the results of research that was carried out to develop highly effective parallel methods of multicriteria optimization making use of all the search information obtained in the course of calculations \cite{c19,c20,c21}.
-A new contribution to this research area consists in the development of an approach for solving MCO problems, in which some of the variable parameters can only take discrete values.
+A new contribution to this research area consists in the development of an approach for solving MCO problems, in which some of the varied parameters can only take discrete values.
 The proposed approach reduces the solution of mixed-integer optimization problems to a family of optimization problems where only continuous parameters are used.
-All problems of the family are solved simultaneously in the time division mode, where the selection of the optimization problem for the next iteration of the global search is performed adaptively taking into account the search information obtained in the course of the calculations.
+All problems of the family are solved simultaneously in the time-shared mode, where the selection of the optimization problem for the next iteration of the global search is performed adaptively taking into account the search information obtained in the course of the calculations.
 The algorithms developed enable efficient parallel computing on high-performance computing systems.
@@
 This section also describes a dimensionality reduction scheme, which allows multidimensional optimization problems to be reduced to a one-dimensional global search problem.
-Section \ref{sec:4} presents a parallel algorithm for solving multicriteria multi-integer optimization problems, which is used in the framework of the proposed approach.
+Section \ref{sec:4} presents a parallel algorithm for solving multicriteria mixed-integer optimization problems, which is used in the framework of the proposed approach.
 Section \ref{sec:5} contains the results of numerical experiments confirming that the proposed approach is quite promising.
@@
 \end{equation}
-where $f(y,u)=(f_1 (y,u),f_2 (y,u), \dots,f_s (y,u))$ is the vector criterion of efficiency in which the variable parameters belong to two different types:
+where $f(y,u)=(f_1 (y,u),f_2 (y,u), \dots,f_s (y,u))$ is the vector criterion of efficiency in which the varied parameters belong to two different types:
 \begin{itemize}
@@
 \end{equation}
-where $U_i=\{v_{i1},v_{i2}, \dots ,v_{il_i} \}$, $1 \leq i \leq m$, is a set of $l_i>0$ admissible discrete values for the parameter $u_k$, i.e., the whole set of all possible discrete parameter values $U$ contains
+where $U_i=\{v_{i1},v_{i2}, \dots ,v_{il_i} \}$, $1 \leq i \leq m$, is a set of $l_i>0$ admissible discrete values for the parameter $u_i$, i.e., the whole set of all possible discrete parameter values $U$ contains
 \begin{equation}\label{eq:4}
@@
 
-In the most complex variant, the criteria $f_i(y,u)$, $1 \leq i \leq s$, may be multi-extremal, and the procedures for calculating their values can be computationally time-consuming.
+In the most complex variant, the criteria $f_i(y,u)$, $1 \leq i \leq s$, may be multiextremal, and the procedures for calculating their values can be computationally time-consuming.
 It is also assumed that the criteria $f_i (y,u)$, $1 \leq i \leq s$, meet the Lipschitz condition
@@
 
-Efficiency criteria in an MCO problem are usually controversial and the parameters $(y^*,u^*) \in D \times U$ with the best values for all criteria at the same time may not be available.
+Efficiency criteria in an MCO problem are usually controversial and the parameters $(y^*,u^*) \in D \times U$ with the optimal values for all criteria at the same time may not be available.
 In such situations, it is a common approach for MCO problems to find efficient (non-dominated) decisions, for which an improvement in the values of some criteria leads to a deterioration of efficiency indicators for other criteria.
@@
 
-A commonly used approach to obtaining individual efficient decisions is to transform a vector criterion into some common scalar efficiency index\footnote{It should be noted that this approach ensures that a wide range of already existing global optimization methods can be used to solve MCO problems.}
+A commonly used approach to obtaining efficient decisions is to transform a vector criterion into some combined scalar efficiency function\footnote{It should be noted that this approach ensures that a wide range of already existing global optimization methods can be used to solve MCO problems.}
 \begin{equation}\label{eq:6}
@@
 \end{equation}
-where $F$ is an objective function generated by scalarization of the criteria $f_i$, $1 \leq i \leq s$, $\alpha$ is the vector of the parameters of the criteria convolution applied, while $D$ and $U$ are the domains of possible parameter values from (\ref{eq:2}-\ref{eq:3}).
+where $F$ is an objective function generated by scalarization of the criteria $f_i$, $1 \leq i \leq s$, $\alpha$ is the vector of the parameters of the criteria convolution applied, while $D$ and $U$ are the domains of possible parameter values from (\ref{eq:2})-(\ref{eq:3}).
 By virtue of (\ref{eq:5}), the function $F(\alpha, y, u)$ also satisfies the Lipschitz condition with some constant $L$, i.e.
@@
 
-To construct a common scalar efficiency index $F(\alpha,y,u)$ from (\ref{eq:6}), one of the most frequently used scalarization methods is to use a minimax criteria convolution \cite{c2,c5}:
+To construct a combined scalar efficiency index $F(\alpha,y,u)$ from (\ref{eq:6}), one of the most frequently used scalarization methods is to use a minimax criteria convolution \cite{c2,c5}:
 \begin{equation}\label{eq:8}
@@
 which is necessary for solving the MCOmix problem.
-This set of problems can be formed sequentially in the course of calculations; the problems of the set can be solved strictly sequentially or simultaneously in time division mode.
+This set of problems can be formed sequentially in the course of calculations; the problems of the set can be solved strictly sequentially or simultaneously in time-shared mode.
 In addition, the problems of the set $\mathbb{F}_T$ can be solved in parallel using high-performance computing systems.
@@
 \begin{matrix}
-  F(\alpha,y^*,u^*)=\min{F(\alpha,y,u)} \min_{y \in D} \min_{u \in U}{F(\alpha,y,u)} =\\ 
-  \min_{y \in D}(F(\alpha,y,w_1 ),F(\alpha,y,w_2 ), \dots, F(\alpha,y,w_l )),w_i\in U,1 \leq i \leq l.
+  F(\alpha,y^*,u^*)=\min{F(\alpha,y,u)} = \min_{y \in D} \min_{u \in U}{F(\alpha,y,u)} = \\ 
+  \min_{y \in D}(F(\alpha,y,w_1 ),F(\alpha,y,w_2 ), \dots, F(\alpha,y,w_l )),\\
+	w_i\in U,1 \leq i \leq l.
 \end{matrix}
@@
 \end{equation}
-having one continuous and one discrete parameter, the plots of the  function $F(\alpha,y,u)$ for different values of the discrete parameter have \hl{form from }Fig.~\ref{fig:1} 
+having one continuous and one discrete parameter, the plots of the  function $F(\alpha,y,u)$ for different values of the discrete parameter have form from Fig.~\ref{fig:1} 
 \begin{figure}
@@
   \includegraphics[width=0.9\linewidth]{fig1}
-  \caption{The function $F(\alpha,y,u)$ of the problem (\ref{eq:12}) plotted for different values of the discrete parameter (left, $u=1$,  right, $u=1$)}
+  \caption{The function $F(\alpha,y,u)$ of the problem (\ref{eq:12}) plotted for different values of the discrete parameter (left, $u=1$,  right, $u=2$)}
   \label{fig:1}
@@
 In some rare situations, the information about the values of the discrete parameters that do not allow the minimum value of the function $F(\alpha,y,u)$ to be reached, may be known a priori.
-In most cases, such information can be obtained only in the process of calculations based on the obtained search information.
+In most cases, such information can be obtained only in the process of calculations based on the calculated search information.
 
@@
 \end{equation}
-is the search information obtained in the process of calculation after global search iterations.
+is the search information obtained in the process of calculation after $k>1$ global search iterations.
 Then the procedure for adaptive estimation of the discrete parameter value, at which the minimum value  of the function $F(\alpha,y,u)$ is expected to be obtained, can be determined using the decision rule
@@
 
-With the decision rule $\Theta$, the general computational scheme for solving the problem can be presented as follows.
+With the decision rule $\Theta$ from (\ref{eq:14}), the general computational scheme for solving the problem can be presented as follows.
 
@@
 \textit{Rule 3.} Check the stopping condition for the computation.
-If the required accuracy of the global search is not achieved, it is necessary to supplement the search information $\omega_k$ from (\ref{eq:13}) with the results of the current iteration and execute rule 1.
+If the required accuracy of the global search is not achieved, it is necessary to supplement the search information $\omega_k$ from (\ref{eq:13}) with the results of the current iteration and continue calculations from Rule 1.
 
@@
 The problem of increasing computational complexity was even called the ``dimensionality curse''.
-In particular, when applying the computational scheme discussed in paragraph \ref{subsec:31}, it is necessary to accumulate and analyze multidimensional search information $\omega_k$ from (\ref{eq:13}).
+In particular, when applying the computational scheme discussed in subsection \ref{subsec:31}, it is necessary to accumulate and analyze multidimensional search information $\omega_k$ from (\ref{eq:13}).
 This computational complexity can be significantly reduced by reducing the optimization problems to be solved by using Peano \textit{curves} or \textit{evolvents} $y(x)$, which unmbiguously and continuously map a segment [0,1] on an $n$-dimensional hyper-cube $D$ -- see, for example, \cite{c24,c25}.
-As a result of this reduction, the multidimensional global optimization problem (\ref{eq:6}) is reduced to a one-dimensional problem:
+As a result of this reduction, the multidimensional global optimization problem (\ref{eq:6}) is reduced to a one-dimensional problem
 \begin{equation}\label{eq:15}
@@
 
-It should be noted that the one-dimensional functions $F(\alpha,y(x),u)$ resulting from the reduction, with a fixed value of discrete parameters $u \in U$ satisfy the uniform H{\" o}lder condition, i.e.
+It should be noted that the one-dimensional functions $F(\alpha,y(x),u)$ resulting from the reduction satisfy the uniform H{\" o}lder condition, i.e.
 \begin{equation}\label{eq:16}
@@
 \end{equation}
-where the constant $H$ is determined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:7}) and $n$ is the dimensionality of the optimization problem (\ref{eq:1}).
+where the constant $H$ is determined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:7}) and $n$ is the dimensionality of the MCOmix problem from (\ref{eq:1}).
 
@@
 
-The use of dimensionality reduction makes it possible to combine (concatenate) one-dimensional functions into a single one-dimensional function $\Phi(\alpha,y(x))$, defined over the segment $[0,l]$ (see Fig.~\ref{fig:1}).
+The use of dimensionality reduction makes it possible to combine (concatenate) one-dimensional functions $F(a, y(x),w_i)$, $1 \leq i \leq l$, from (\ref{eq:15}) into a single one-dimensional function $\Phi(\alpha,y(x))$, defined over the segment $[0,l]$ (see Fig.~\ref{fig:1}).
 
@@
 F(\alpha,y(x),w_1 ),x \in [0,1], \\
-F(\alpha,y(x),w_2 ),x \in [1,2], \\
+F(\alpha,y(x),w_2 ),x \in (1,2], \\
 \dots \\
-F(\alpha,y(x),w_l ),x \in [l-1,l],
+F(\alpha,y(x),w_l ),x \in (l-1,l],
 \end{cases}
@@
 ($E(x)$ denotes the operation of taking the integral part of the number $x$).
-It should be noted that the function $\Phi(\alpha,y(x))$ is discontinuous at the points $x_i$, $1 \leq i \leq l-1$,  further on, the values of the function $\Phi(\alpha,y(x))$ at these points are considered undefined are not used in the calculations.
+It should be noted that the function $\Phi(\alpha,y(x))$ is discontinuous at the points $x_i$, $1 \leq i \leq l-1$,  further on, the values of the function $\Phi(\alpha,y(x))$ at these points are considered undefined and not used in the calculations.
 
@@
 The proposed approach uses the algorithm of global mixed-integer search (AGMIS) to minimize the function $F(\lambda,y,u)$.
-This algorithm expands the possibilities of multi-extremal optimization methods developed within the framework of the information-statistical theory of global search \cite{c19,c20,c21,c22,c23,c32,c33,c34,c35,c36,c37,c38,c39} to minimize the reduced one-dimensional function $\Phi(\alpha,y(x))$ from (\ref{eq:19}).
+This algorithm expands the possibilities of multiextremal optimization methods developed within the framework of the information-statistical theory of global search \cite{c19,c20,c21,c22,c23,c32,c33,c34,c35,c36,c37,c38,c39} to minimize the reduced one-dimensional function $\Phi(\alpha,y(x))$ from (\ref{eq:19}).
 
@@
 Next, let $k$, $k > 1$, global search iterations be performed.
-Selection of the test points $(k+1)$ for the next iteration is determined by the following rules.
+Selection of the trial points $(k+1)$ for the next iteration is determined by the following rules.
 
@@
 \end{equation}
-(the values of the discrete parameters u are given according to (\ref{eq:19})).
+(the values of the discrete parameters $u \in U$ are given according to (\ref{eq:19})).
 
@@
 The calculated characteristics $R(i)$, $1<i \leq k$, can be interpreted as some measure of the importance of intervals in terms of the presence of the global minimum point.
-It explains the interval selection scheme for the next trial -- the point of each subsequent trial $x^{k+1}$ from (\ref{eq:22}) is chosen in the range with the maximum value of the interval characteristic (i.e., in the interval  where the global minimum is most likely to be located).
+The interval selection scheme for the next trial becomes clear -- the point of each subsequent trial $x^{k+1}$ from (\ref{eq:22}) is chosen in the interval with the maximum value of the interval characteristic (i.e., in the interval  where the global minimum is most likely to be located).
 
-It should also be noted that the AGMIS computational scheme considered above refines the general scheme of global search for mixed-integer optimization problems from paragraph \ref{subsec:31}.
+It should also be noted that the AGMIS computational scheme considered above refines the general scheme of global search for mixed-integer optimization problems from subsection \ref{subsec:31}.
 Thus, the choice of the function $F(\alpha, y(x), w_i)$, $1\leq i\leq l$, for the next iteration is provided by the procedure of finding the interval with the maximum characteristic.
 
-A full description of multi-extremal optimization algorithms and conditions for their convergence developed within the framework of the information-statistical theory of global search are given in \cite{c24}.
+A full description of multiextremal optimization algorithms and conditions for their convergence developed within the framework of the information-statistical theory of global search are given in \cite{c24}.
 Thus, with a proper numerical estimation of the H{\"o}lder constants $H$ from (\ref{eq:16}), the AGMIS algorithm converges to all available points of the global minimum of the minimized function $F(\lambda, y, u)$.
@@
 Now, let us return to the initial problem statement (\ref{eq:1}) and remind that for solving the MCOmix problem it may be necessary to solve the set of problems $\mathbb{F}_T$ from (\ref{eq:10}).
-In solving this set of problems with the help of the AGMIS algorithm, one more key property of the proposed approach is manifested -- the results of all earlier  calculations of criteria values can be carried into the values of the next optimization problem $F(\alpha, y(x), u)$ from (\ref{eq:6}) to be solved with new values of $\alpha'$ from (\ref{eq:8}) without repeating any time-consuming calculations of criteria values, i.e.
+In solving this set of problems by using the AGMIS algorithm, one more key property of the proposed approach is manifested -- the results of all previous  calculations of criteria values can be recalculated into the values of the next optimization problem $F(\alpha, y(x), u)$ from (\ref{eq:6}) to be solved with new values of $\alpha'$ from (\ref{eq:8}) without repeating any time-consuming calculations of criteria values, i.e.
 \begin{equation}\label{eq:24}
@@
 
-The final stage in improving the efficiency of solving MCOmix problems in the proposed approach is to organize parallel computing on modern high-performance systems.
+The final stage in improving the efficiency of solving MCOmix problems in the proposed approach is to organize parallel computing on parallel high-performance systems.
 Unfortunately, attempts to develop parallel variants of AGMIS and AMMIS algorithms using existing parallelization methods have not succeeded.
@@
 \textit{Rule 3 $'$}.
-Perform new trials (calculate the values of the minimized function $F(\alpha,y(x),u)$ at the points $x^{k+j}$, $1 \leq j \leq p$ located in the intervals with the maximum characteristics from (\ref{eq:25}).
+Perform new trials (calculate the values of the minimized function $F(\alpha,y(x),u)$) at the points $x^{k+j}$, $1 \leq j \leq p$ located in the intervals with the maximum characteristics from (\ref{eq:25}).
 
@@
 
-For AMMIS, 50 global optimization problems with different sets of convolution coefficients were solved.
+For AMMIS, 50 global optimization problems with different values of convolution coefficients $\lambda\in \Lambda$ from (\ref{eq:8}) were solved.
 The accuracy of the method $\varepsilon=0.05$ and the reliability parameter $r=3.0$ from (\ref{eq:23}) were used when solving the series of problems.
-The results of the experiments are presented in full in Table \ref{tab:1}.
+The results of the experiments are presented in Table \ref{tab:1}.
 
@@
 To solve each MCOmix problem, 50 uniformly distributed values of convolution coefficients $\lambda \in \Lambda$ from (\ref{eq:8}) were used.
-To estimate the values of HV and DU efficiency indicators, an approximation of the Pareto domain was calculated for each MCOmix problem  using a uniform grid in the domain of variable values $D\times U$ from (\ref{eq:2})-(\ref{eq:3}).
+To estimate the values of HV and DU efficiency indicators, an approximation of the Pareto domain was calculated for each MCOmix problem  using a uniform grid in the domain of varied values $D\times U$ from (\ref{eq:2})-(\ref{eq:3}).
 The PAMMIS algorithm was used for the following parameter values: accuracy $\varepsilon=0.05$ from (\ref{eq:26}) and reliability $r=5.6$ (the $r$ parameter is used in PAMMIS to construct estimates of the H{\"o}lder constant $H$ from (\ref{eq:16})).
@@
 \label{tab:2}
-\begin{tabular}{cccccc}
+\begin{tabular}{ccccccc}
 \hline
-Number of computing cores    & Number of iterations    & Speedup  & DU index&   & HV index  \\\hline
-\multicolumn{6}{c}{The   Pareto domain estimate obtained by exhaustive search}        \\\hline
-48                           & 1 706 667               & -        & 20,6 & & 29 784,9  \\\hline
-\multicolumn{6}{c}{The Pareto    domain  estimate obtained using the   PAMMIS method} \\\hline
-1                            & 86 261,6                & 1,0      & 13,8 & & 30 212,3  \\
-6                            & 15 406,2                & 5,6      & 14,4 & & 30 317,6  \\
-12                           & 6 726,1                 & 12,8     & 16,2 & & 30 248,3  \\
-18                           & 5 972,7                 & 14,4     & 14,3 & & 30 551,1  \\
-24                           & 3 657,4                 & 23,6     & 13,9 & & 30 092,3  \\
-30                           & 3 162,3                 & 27,3     & 14,9 & & 30 443,9  \\
-36                           & 2 888,7                 & 29,9     & 14,6 & & 30 528,7  \\
-42                           & 2 028,8                 & 42,5     & 15,6 & & 30 046,1  \\
-48                           & 1 767,6                 & 48,8     & 15,3 & & 30 255,5 \\ \hline
+Cores                  & Iterations                  & Speedup                 & DU           & &       & HV                      \\ \hline
+\multicolumn{7}{c}{\begin{tabular}[c]{@{}c@{}}The Pareto domain estimate obtained \\ by exhaustive search\end{tabular}}        \\ \hline
+48                     & 1 706 667                   & -                       & 20,6         & &       & 29 784,9                \\ \hline
+\multicolumn{7}{c}{\begin{tabular}[c]{@{}c@{}}The Pareto    domain  estimate obtained \\ using the PAMMIS method\end{tabular}} \\ \hline
+1                      & 86 261,6                    & 1,0                     & 13,8         & &      & 30 212,3                \\
+6                      & 15 406,2                    & 5,6                     & 14,4         & &      & 30 317,6                \\
+12                     & 6 726,1                     & 12,8                    & 16,2         & &      & 30 248,3                \\
+18                     & 5 972,7                     & 14,4                    & 14,3         & &      & 30 551,1                \\
+24                     & 3 657,4                     & 23,6                    & 13,9         & &      & 30 092,3                \\
+30                     & 3 162,3                     & 27,3                    & 14,9         & &      & 30 443,9                \\
+36                     & 2 888,7                     & 29,9                    & 14,6         & &      & 30 528,7                \\
+42                     & 2 028,8                     & 42,5                    & 15,6         & &      & 30 046,1                \\
+48                     & 1 767,6                     & 48,8                    & 15,3         & &      & 30 255,5                \\ \hline
 \end{tabular}
@@
 The results of the computational experiments show that the PAMMIS algorithm is scalable as the  speedup of parallel computations increases almost linearly with the number of computing cores used.
-The obtained values of HV and DU indicators show that the Pareto set estimate using the PAMMIS algorithm is calculated with a greater accuracy and requires a significantly smaller number of trials compared to the results obtained using uniform grids in the domain of variable values $D \times U$.
+The obtained values of HV and DU indicators show that the Pareto set estimate using the PAMMIS algorithm is calculated with a greater accuracy and requires a significantly smaller number of trials compared to the results obtained using uniform grids in the domain $D \times U$.
 
@@
 
-In this paper, a new approach is proposed to solving computationally intensive optimization problems, where some variable parameters can only take on discrete values (MCOmix).
-It is also assumed that the optimality criteria may be multiextremal, and the computing of criterion values may involve a large amount of calculations.
+In this paper, a new approach is proposed to solving computationally intensive optimization problems, where some varied parameters can only take on discrete values (MCOmix).
+It is also assumed that the efficiency criteria may be multiextremal, and the computing of criterion values may involve a large amount of calculations.
 Due to the high computational complexity of this class of problems, parallel solution methods need to be developed, relying on the efficient use of high-performance computing systems.
@@
 In the framework of the proposed approach, solution of mixed-integer optimization problems is reduced to solving a family of global search optimization problems that use only continuous parameters.
-All problems of the family are solved simultaneously in the time division mode, where the optimization problem for the next global search iteration is selected adaptively taking into account the search information obtained in the course of computations.
-The results of numerical experiments prove that  this approach can significantly reduce computational intensity when solving MCOmix problems in a multistage procedure.
+All problems of the family are solved simultaneously in the time-shared mode, where the optimization problem for the next global search iteration is selected adaptively taking into account the search information obtained in the course of computations.
+The results of numerical experiments prove that  this approach can significantly reduce computational intensity when solving MCOmix problems.
 
-In conclusion, it may be noted that the proposed approach is quite promising and requires continued investigations.
-First of all, it is necessary to continue numerical experiments to solve multicriteria mixed optimization problems with a greater number of efficiency criteria and having a greater dimensionality.
+In conclusion, it may be noted that the proposed approach is promising and requires continued investigations.
+First of all, it is necessary to continue numerical experiments to solve multicriteria mixed-integer optimization problems with a greater number of efficiency criteria and having a greater dimensionality.
 The possibility of organizing parallel computations using high-performance systems with distributed memory should be also evaluated.

### 2020-12-13 2021/PСT_GergelKozinov/2021_PCT.tex
2021/PСT_GergelKozinov/2021_PCT.tex
Исправление ошибок в статье.

@@
 Section \ref{sec:4} presents a parallel algorithm for solving multicriteria mixed-integer optimization problems, which is used in the framework of the proposed approach.
-Section \ref{sec:5} contains the results of numerical experiments confirming that the proposed approach is quite promising.
+Section \ref{sec:5} contains the results of numerical experiments confirming that the proposed approach is promising.
 In conclusion, the results obtained are discussed and possible main directions for continuing the research are outlined.
@@
 
-To construct a combined scalar efficiency index $F(\alpha,y,u)$ from (\ref{eq:6}), one of the most frequently used scalarization methods is to use a minimax criteria convolution \cite{c2,c5}:
+To construct a combined scalar efficiency function $F(\alpha,y,u)$ from (\ref{eq:6}), one of the most frequently used scalarization methods is to use a minimax criteria convolution \cite{c2,c5}:
 \begin{equation}\label{eq:8}
@@
 
-The use of dimensionality reduction makes it possible to combine (concatenate) one-dimensional functions $F(a, y(x),w_i)$, $1 \leq i \leq l$, from (\ref{eq:15}) into a single one-dimensional function $\Phi(\alpha,y(x))$, defined over the segment $[0,l]$ (see Fig.~\ref{fig:1}).
+The use of dimensionality reduction makes it possible to combine (concatenate) one-dimensional functions $F(\alpha, y(x),w_i)$, $1 \leq i \leq l$, from (\ref{eq:15}) into a single one-dimensional function $\Phi(\alpha,y(x))$, defined over the segment $[0,l]$ (see Fig.~\ref{fig:1}).
 
@@
 \section*{Acknowledgements} 
-\hl{This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''}
+This work was supported by the project no. 0729-2020-0055 and by the Research and Education Mathematical Center, project no. 075-02-2020-1483/1
 

### 2020-12-14 2021/ICCS_BarkalovUsova/iccs.tex
Added bibliography, formules, figures and tables

@@
 \usepackage{graphicx}
+\usepackage{enumitem}
+\usepackage{amsmath,amssymb}
+\usepackage{mathtext}
 
-%%%% ДЛЯ РУССКОГО ТЕКСТА
+\newcommand{\sign}{\operatorname{sign}}
+
+%%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
 %\usepackage[utf8x]{inputenc}
@@
 Numerical methods.}
+
 \end{abstract}
@@
 
-\section{The global search algorithm}
-\label{SectionCore}
+$x^*$
 
-\section{Results of numerical experiments}
+$\varphi(x)$
 
-Example of references 
-The comparison of the algorithms has been carried out using the Grishagin test problems $F_{gr}$ (see, for example, \cite{Grishagin1994}, test function 4) and the GKLS generator \cite{Gaviano2003}.
-In \cite{Barkalov2015,Grishagin2018} the global search algorithm (GSA) with the use of the evolvents as well as in combination 
-with the adaptive dimensionality reduction scheme was shown to overcome many well known optimization 
-algorithms including DIRECT \cite{Jones1993} and DIRECT\textit{l} \cite{Gablonsky2001}.
-Therefore, in the present study we will limit ourselves to the comparison of the  variants of GSA with different dimensionality reduction schemes.
+\begin{equation}\label{problem}
+\varphi(x^*)=\min\left\{\varphi(x):x\in\left[a,b\right]\right\}.
+\end{equation}
+
+(see, for example, popular approaches in \cite{Sergeyev2013}, \cite{PaulaviciusZilinskas2014}, \cite{Sergeyev2017}).
+
+(\ref{problem})
+
+\cite{Horst1995}, \cite{Horst1996}, \cite{Pinter1996})
+
+$\varphi(x)$
+
+$\varphi(x)$  $L$
+
+\[
+\left|\varphi(x')-\varphi(x'')\right|\leq L\left|x'-x''\right|,\; x',x'' \in [a,b].
+\]
+
+\cite{Jones2009}, \cite{Evtushenko2009}, \cite{Paulavicius2011}, \cite{Evtushenko2013}.
+
+\begin{itemize}
+  \item
+  \item 
+  \item
+\end{itemize}
+
+\cite{Moreau2000}, \cite{Batukhtin1993}, \cite{Batukhtin1998}
+\cite{Ban2019}, \cite{ZhangXu}
+
+\cite{Strongin2000}
+
+ MATLAB Global Optimization Toolbox \cite{MatlabOTB}. Section 6 concludes the paper.
+
+\section{2}
+
+(\ref{problem})
+
+objective function $\varphi(x):x\in[a,b]$
+
+\begin{equation}\label{points}
+a = \omega_0 <\omega_1 < ... <\omega_s < \omega_{s+1} = b.
+\end{equation}
+
+\begin{equation}\label{LipschitzСondition}
+\left|\varphi(x')-\varphi(x'')\right|\leq L\left|x'-x''\right|,\; x',x'' \in (\omega_{i-1},\omega_{i}), 1 \leq i \leq s+1.
+\end{equation}
+
+$\omega_{0}$ and $\omega_{s+1}$
+
+$\Omega$ $\Omega = \{\omega_{0}, ..., \omega_{s+1}\} \omega_{0} = a, \omega_{s+1} = b$
+
+\cite{Strongin2000}
+
+$s+1$
+
+$x^i \in (\omega_{i-1},\omega_{i}), 1 \leq i \leq s+1 $
+
+$x^{k+1}$, $k \geq s+1$
+
+$x^1,…,x^k$
+
+$\omega_{0}, ..., \omega_{s+1}$
+
+$x_i, 0\leq i \leq m = k + s + 1$
+
+\begin{equation}\label{pointsX}
+a = x_0 < x_1 < ... < x_{m-1} < x_{m} = b,
+\end{equation}
+
+$z_i=\varphi(x_i)$
+
+$x_i \not\in \Omega$
+
+$(x_{i-1},x_i )$
+
+$(x_{i-1},x_i ), 1\leq i \leq m$
+
+\begin{equation}\label{mu_i}
+\mu_i=(1-\gamma_i-\gamma_{i-1} )  \frac{|z_i-z_{i-1} |}{(x_i-x_{i-1} )}
+\end{equation}
+
+\begin{equation}\label{gamma}
+\gamma_i = 
+\begin{cases}
+	1, &\text{$x_i \in \{\omega_{0},...,\omega_{s+1}\} $} \\
+	0, &\text{$x_i \not\in \{\omega_{0},...,\omega_{s+1}\} $}
+\end{cases}
+\end{equation}
+
+\begin{equation}\label{mu}
+\mu=\max\left\{\mu_i: 1 \leq i \leq m\}\right\}
+\end{equation}
+
+$\mu=0$ $\mu=1$
+
+$ 0 \leq \gamma_i+\gamma_{i-1} \leq 1$
+
+$ 1 \leq i \leq m$
+
+ $\mu_i$
+
+ $0$
+
+$ (x_{i-1},x_i ),1 \leq i \leq m$
+
+$\Omega$
+
+$(x_{i-1},x_i ), 1\leq i \leq m$
+
+\begin{equation}\label{R_i}
+R(i)=(1+\gamma_i+\gamma_{i-1} )\Delta_i+(1-\gamma_i-\gamma_{i-1} ) \frac{ (z_i-z_{i-1} )^2}{(r\mu)^2 \Delta_i }  -\frac{2[(1-\gamma_i )(1+\gamma_{i-1} ) z_i+(1+\gamma_i )(1-\gamma_{i-1} ) z_{i-1} ]}{r\mu}
+\end{equation}
+
+$r>1$
+
+\begin{equation}\label{R}
+R(t)=\max\left\{R(i): 1 \leq i \leq m\}\right\}
+\end{equation}
+
+\begin{equation}\label{new_x}
+x^{k+1}=\frac {(x_t+x_{t-1})}{2}-(1-\gamma_t-\gamma_{t-1} ) \frac{(z_t-z_{t-1})}{2r\mu}
+\end{equation}
+
+$x_t-x_{t-1} \leq \epsilon$
+
+$\epsilon > 0$
+
+\begin{equation}\label{testFunction}
+\varphi(x) =  0.1\sum_{i=1}^5{i\sin{(10(i+1)x+i)}}+
+\begin{cases}
+	-4, &0 \leq x < 3.1\\
+	10, &3.1 \leq x < 4.6\\
+	0, &4.6 \leq x < 7.0\\
+	30, &7.0 \leq x < 9.0\\
+	0, &9.0 \leq x \leq 10.0
+\end{cases}
+\end{equation}
+
+$x\in[0,10]$
+
+$r=2.0,\epsilon=0.001$
+
+\begin{figure}[!ht]
+	\begin{center}
+			\includegraphics[width=0.9\linewidth]{ris_1.jpg}
+			\caption{Fig.1}
+      \label{ris1}
+	\end{center}
+\end{figure}
+
+\section{3}
+
+$\varphi(x)$
+
+$\omega \in(x_{i-1},x_i)$
+
+$\mu_i$
+
+$\mu_i$
+
+$(x_{i-1},x_i)$
+
+$\mu_i$
+
+\begin{equation}\label{mu_cond}
+\mu(1) \leq \mu(2) \leq ... \leq \mu(m)
+\end{equation}
+
+\begin{equation}\label{mu_main_cond}
+ \frac{\mu(p)}{\mu(p+1)} \leq Q, 1 \leq p < q(m-2(s+1))
+\end{equation}
+
+$Q >1$ and $0<q<1$
+
+$\mu_i$
+
+$p<m-2(s+1)$
+
+$1 \leq p<q(m-2(s+1))$
+
+$\mu_i$
+
+$I=\{i:1 \leq i \leq m,\mu_i=\mu(j),1 \leq j \leq p\}$
+
+$(x_{i-1},x_i)$
+
+$\mu_i$
+
+$(x_{i-1},x_i), i \in I$
+
+\section{4}
+
+$x^i \in (\omega_{i-1},\omega_i ),1 \leq i \leq s+1$
+
+$\Omega=\{\omega_0,…,\omega_{s+1}\}$
+
+$\omega_0=a$ and $\omega_{s+1}=b$
+
+$x^{k+1}, k \geq s+1$
+
+$\{x^1,...,x^k\} \cup \Omega$
+
+\begin{equation}\label{x_cond}
+ a=x_0<x_1<...<x_{m-1}<x_m=b,
+\end{equation}
+
+$m=s+k+1$
+
+$\gamma_i$
+
+$z_i=\varphi(x_i )$
+
+$ x_i \not\in \Omega$
+
+$\gamma_i=0$
+
+$\mu_i,1 \leq i \leq m$
+
+$0<q<1<Q$
+
+\begin{equation}\label{delta_cond}
+\delta_i=
+\begin{cases}
+	\sign{(z_i - z_{i-1})}, &\text{$i \in I$} \\
+	0, &\text{$i \not \in I$}
+\end{cases}
+\end{equation}
+
+$\delta_i=-1$
+
+$(x_{i-1},x_i ), \delta_i=1$
+
+$\delta_i=0$
+
+$ 0 \leq \gamma_i+\gamma_{i-1}+|\delta_i | \leq 1$
+
+$1 \leq i \leq m$
+
+\begin{equation}\label{mu_new}
+\mu=\max\left\{(1-\gamma_i-\gamma_{i-1}-|\delta_i|) \frac {|z_i-z_{i-1}|}{(x_i-x_{i-1})}: 1 \leq i \leq m\}\right\}
+\end{equation}
+
+$\mu=0$ $\mu=1$
+
+$(x_{i-1},x_i), 1 \leq i \leq m$
+
+\begin{equation}\label{mu_new}
+\begin{gathered}
+R(i)=(1+\gamma_i+\gamma_(i-1)+|\delta_i |) \Delta_i+(1-\gamma_i-\gamma_{i-1}-|\delta_i |)\frac{(z_i-z_{i-1} )^2}{(r\mu)^2 \Delta_i } \\
+- \frac{2[((1-\gamma_i )(1+\gamma_{i-1} )(1-\delta_i )) z_i+((1+\gamma_i )(1-\gamma_{i-1} )(1+\delta_i )) z_{i-1} ]}{r\mu}
+\end{gathered}
+\end{equation}
+
+$\mu$ $\Delta_i=x_i-x_{i-1}$
+
+\begin{equation}\label{R}
+R(t)=\max\left\{R(i): 1 \leq i \leq m\}\right\}
+\end{equation}
+
+\begin{equation}\label{new_x_2}
+x^{k+1}=\frac {(x_t+x_{t-1})}{2}-(1-\gamma_t-\gamma_{t-1}- |\delta_t|) \frac{(z_t-z_{t-1})}{2r\mu}
+\end{equation}
+
+$x_t-x_{t-1} \leq \epsilon$
+
+$\epsilon > 0$
+
+$r=2.0,\epsilon=0.001,Q=3.0,q=0.3$
+
+\begin{figure}[!ht]
+	\begin{center}
+			\includegraphics[width=0.9\linewidth]{ris_2.jpg}
+			\caption{Fig.2}
+      \label{ris2}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+	\begin{center}
+			\includegraphics[width=0.9\linewidth]{ris_3.jpg}
+			\caption{Fig.3}
+      \label{ris3}
+	\end{center}
+\end{figure}
+
+$r=2.0,\epsilon=0.001,Q=3.0,q=0.3$
+
+\section{Results of Numerical Experiments}
+
+\cite{Audet}
+\cite{Goldberg}
+\cite{Kirkpatrick}
+\cite{MatlabOTB}
+
+\begin{equation}\label{seriaFunction}
+f(x) =  \sum_{i=0}^4{(j+1)A_j\sin{(5\pi jx+j)}+(j+1)B_j\cos(3\pi jx+j)}, x\in[0,1]
+\end{equation}
+
+$A_j,B_j$
+
+$[-6,6]$
+
+$\omega_1,\omega_2,\omega_3,\omega_4$
+
+$[0,0.25), [0.25,0.5), [0.5,0.75), [0.75,1]$
+
+$\delta_1,\delta_2,\delta_3,\delta_4$
+
+$[-50,50]$
+
+\begin{equation}\label{seriaFunction_jump}
+\varphi(x) =  f(x)+
+\begin{cases}
+	\delta_1, &0 \leq x < \omega_1\\
+	\delta_2, &\omega_1 \leq x < \omega_2\\
+	\delta_3, &\omega_2 \leq x < \omega_3\\
+	\delta_4, &\omega_3 \leq x < \omega_4\\
+	\delta_1, &\omega_4 \leq x \leq 1.0
+\end{cases}
+\end{equation}
+
+$x\in[0,1]$
+
+global optimizer $x^*$
+
+$x^k$
+
+$|x^k-x^* | \leq \delta$
+
+$x^*$
+
+$\delta= 10^{-2}$
+
+$\epsilon = 10^{-3}$
+
+$x_0$
+
+ MATLAB Global Optimization Toolbox
+
+$r=2.2$
+
+\begin{table}[h]
+	\caption{Results}
+	\begin{center}
+		\begin{tabular}{|c|c|c|c|}
+			\hline
+			Method & Solved & Sr count of iteration & Sr funccount \\
+			\hline
+			\hline
+                                AGS-D & 1000  &  52  &  53 \\
+                                \hline
+                                SA    &  1000  &  774  &  777  \\
+			\hline
+			GA    &   999  &  24  &  1290  \\
+			\hline
+                                DS    &  970  & 38  &  71  \\
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
+$r=3.7,Q=3.0,q=0.3$
+
+\begin{table}[h]
+	\caption{Results 2}
+	\begin{center}
+		\begin{tabular}{|c|c|c|c|}
+			\hline
+			Method & Solved & Sr count of iteration & Sr funccount \\
+			\hline
+			\hline
+                                AGS-D & 1000  & 79  &  80 \\
+                                \hline
+                                SA    &  998  &  765  &  770  \\
+			\hline
+			GA    &   993  &  25  &  1310  \\
+			\hline
+                                DS    &  964  & 38  &  71  \\
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
+Word order:
+
+\cite{Sergeyev2013}
+\cite{PaulaviciusZilinskas2014}
+\cite{Sergeyev2017}
+\cite{Horst1995}
+\cite{Horst1996}
+\cite{Pinter1996}
+\cite{Jones2009}
+
+\cite{Evtushenko2009}
+\cite{Paulavicius2011}
+\cite{Evtushenko2013}
+\cite{Strongin2000}
+
+\cite{Moreau2000}
+
+\cite{Batukhtin1993}
+\cite{Batukhtin1998}
+\cite{Ban2019}
+
+\cite{ZhangXu}
+\cite{Audet}
+
+\cite{Goldberg}
+\cite{Kirkpatrick}
+\cite{MatlabOTB}
 

### 2020-12-16 2021/ICCS_BarkalovUsova/iccs.tex
main text is ready

--- 2020-12-02 2021/ICCS_BarkalovUsova/iccs.tex
@@
 \usepackage{graphicx}
+\usepackage{enumitem}
+\usepackage{amsmath,amssymb}
+\usepackage{mathtext}
 
-%%%% ДЛЯ РУССКОГО ТЕКСТА
+\usepackage{hyperref}
+
+\newcommand{\sign}{\operatorname{sign}}
+
+%%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
 %\usepackage[utf8x]{inputenc}
@@
 %
-\title{An algorithm for searching the global minimum of univariate discontinuous functions\thanks{This work was supported by project no. 0729-2020-0055.}}
+\title{An algorithm for searching the global minimum of univariate discontinuous functions\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.
+0729-2020-0055, and by the Research and Education Mathematical Center, project no.
+075-02-2020-1483/1.}}
 %
@@
 \begin{abstract}
+We consider the problem of finding the global minimum of a one-dimensional function that may have several finite jump discontinuity points.
+The discontinuities of the objective function can reflect the specifics of the optimized object in the mathematical model (for example, shock effects, resonance phenomena, jumps in geometric dimensions or material properties, etc.).
+In some cases, the discontinuing set is known.
+However, at the same time, there are problems in which there are no a priori estimates of discontinuity points, but it is known that such points are possible.
+The paper gives a description of the global search algorithm for solving the specified class of problems.
+The authors conducted an experimental comparison of the proposed algorithm with the methods implemented in MATLAB Global Optimization Toolbox.
 
 \keywords{Global optimization \and Multiextremal functions \and 
-Dimensionality reduction \and Peano curve \and Nested optimization \and 
-Numerical methods.}
+Discontinuous functions \and Algorithms' comparison }
+
 \end{abstract}
@@
 
-\section{The global search algorithm}
-\label{SectionCore}
+This paper considers the problem of finding the global minimum $x^*$ of a one-dimensional function $\varphi(x)$ of the form 
+\begin{equation}\label{problem}
+\varphi(x^*)=\min\left\{\varphi(x):x\in\left[a,b\right]\right\}.
+\end{equation}
+One-dimensional optimization problems play an important role, because many approaches to solving multidimensional optimization problems are, in one way or another, based on reducing the solution of the original multidimensional problem to solving a series of related one-dimensional optimization problems (see, for example, popular approaches in \cite{PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
 
-\section{Results of numerical experiments}
+The literature describes various methods for solving the problem (\ref{problem}) depending on the assumptions about the properties of the objective function \cite{Horst1995,Horst1996,Pinter1996}.
+One of such assumptions, which is often fulfilled in applied optimization problems, is that the objective function $\varphi(x)$ satisfies the Lipschitz condition
+\[
+\left|\varphi(x')-\varphi(x'')\right|\leq L\left|x'-x''\right|,\; x',x'' \in [a,b],
+\]
+where the constant $L$ is unknown a priori.
 
-Example of references 
-The comparison of the algorithms has been carried out using the Grishagin test problems $F_{gr}$ (see, for example, \cite{Grishagin1994}, test function 4) and the GKLS generator \cite{Gaviano2003}.
-In \cite{Barkalov2015,Grishagin2018} the global search algorithm (GSA) with the use of the evolvents as well as in combination 
-with the adaptive dimensionality reduction scheme was shown to overcome many well known optimization 
-algorithms including DIRECT \cite{Jones1993} and DIRECT\textit{l} \cite{Gablonsky2001}.
-Therefore, in the present study we will limit ourselves to the comparison of the  variants of GSA with different dimensionality reduction schemes.
+This feature of the objective function is commonly used in many approaches to the development of algorithms for multiextremal optimization \cite{Evtushenko2009,Evtushenko2013,Jones2009,Paulavicius2011}.
+
+However, in some applied problems, the Lipschitz property may not be fulfilled due to the presence of discontinuous jumps in the values of the function at some points of the search domain.
+Such jumps can be caused by sharp changes in the characteristics of the optimized object with small changes in its parameters, for example:
+\begin{itemize}
+  \item changes in the thickness of the material (or other geometric parameters of the object);
+  \item changes in material properties (both when different materials are used in one design, and with changes in the properties of the same material as a result of technological operations, for example, welding, etc.);
+  \item changes in external sources of impact (temperature, force, etc.).
+\end{itemize}
+
+We will interpret such sharp changes as jump discontinuities.
+For these cases, the set of points, at which the characteristics of the object jump, is typically known in advance.
+However, at the same time, there are problems in which there are no a priori estimates of discontinuity points, but it is known that such points are possible.
+
+The known methods for solving such problems, as a rule, either generalize the concept of a gradient for discontinuous functions \cite{Batukhtin1993,Batukhtin1998,Moreau2000}, or belong to the class of bio-inspired algorithms \cite{Ban2019,ZhangXu}.
+These methods provide, generally speaking, the search for only a local solution to the problem.
+
+In this paper, we will consider a deterministic algorithm for solving global optimization problems with both given and unspecified discontinuity points.
+This algorithm was developed on the basis of the \textit{global search algorithm} proposed by Strongin \cite{Strongin2000} and provides a search for the global optimum.
+Section 2 describes the algorithm for functions with given discontinuities.
+Section 3 discusses how to identify unspecified discontinuities.
+Section 4 describes how this discontinuity detection method can be used in the original algorithm.
+Section 5 presents the results of a comparison of the developed algorithm with the global search methods implemented in MATLAB Global Optimization Toolbox \cite{MatlabOTB}.
+Section~6 concludes the paper.
+
+\section{Algorithm for functions with given discontinuity points}
+
+Let us consider problem (\ref{problem}) assuming that the objective function $\varphi(x), x\in[a,b],$ has jump discontinuities at the given points
+\begin{equation}\label{points}
+a = \omega_0 <\omega_1 < ... <\omega_s < \omega_{s+1} = b
+\end{equation}
+and satisfies the Lipschitz condition between them, i.e.
+\begin{equation}\label{LipschitzСondition}
+\left|\varphi(x')-\varphi(x'')\right|\leq L\left|x'-x''\right|, \; x',x'' \in (\omega_{i-1},\omega_{i}), \; 1 \leq i \leq s+1.
+\end{equation}
+Points $\omega_{0}$ and $\omega_{s+1}$ are included in the general list regardless of the presence of function discontinuities in them for the convenience of subsequent presentation.
+Let us denote the resulting set of points $\Omega$, i.e.
+$\Omega = \{\omega_{0}, ..., \omega_{s+1}\}$, where $\omega_{0} = a, \; \omega_{s+1} = b$.
+
+In \cite{Strongin2000}, the global search algorithm (GSA) was considered for solving problems without discontinuities.
+The authors also proposed a method for accounting for discontinuities based on a smoothing transformation of the objective function.
+Below we consider an algorithm based on GSA, in which the discontinuous function is minimized without smoothing of the discontinuities, which greatly simplifies the computational scheme of the algorithm.
+
+In the preliminary step of the method, the first $s+1$ trials are carried out at arbitrary interior points $x^i \in (\omega_{i-1},\omega_{i}), \; 1 \leq i \leq s+1 $.
+A point $x^{k+1}$, $k \geq s+1$, for the next trial is selected as follows.
+
+\begin{enumerate}
+\item Arrange prior trial points $x^1,...,x^k$ and points $\omega_{0}, ..., \omega_{s+1}$ in ascending coordinate order.
+Designate the points of the resulting single series $x_i, 0\leq i \leq m = k + s + 1$,
+\begin{equation}\label{pointsX}
+a = x_0 < x_1 < ... < x_{m-1} < x_{m} = b,
+\end{equation}
+and match them with values $z_i=\varphi(x_i)$ for $x_i \not\in \Omega$.
+Note that, in accordance with the choice of trial points at the preliminary step of the algorithm, there will be no interval $(x_{i-1},x_i)$, the boundary points of which simultaneously belong to the set $\Omega$.
+
+\item For each interval $(x_{i-1},x_i ), 1\leq i \leq m$, calculate the values
+\begin{equation}\label{mu_i}
+\mu_i=(1-\gamma_i-\gamma_{i-1} )  \frac{|z_i-z_{i-1} |}{(x_i-x_{i-1} )}
+\end{equation}
+where
+\begin{equation}\label{gamma}
+\gamma_i = 
+\begin{cases}
+	1, &\text{$x_i \in \{\omega_{0},...,\omega_{s+1}\} $}, \\
+	0, &\text{$x_i \not\in \{\omega_{0},...,\omega_{s+1}\}. $}
+\end{cases}
+\end{equation}
+Find the maximum value 
+%\begin{equation}\label{mu}
+\[
+\mu=\max\left\{\mu_i: 1 \leq i \leq m\right\}.
+\]
+%\end{equation}
+When $\mu=0$, use the value $\mu=1$.
+Note that rule (\ref{gamma}) ensures that the inequality $0 \leq \gamma_i+\gamma_{i-1} \leq 1$ holds for every $1 \leq i \leq m$.
+Therefore, the value $\mu_i$ will be equal to 0 if one of the endpoints of the interval $ (x_{i-1},x_i ), \; 1 \leq i \leq m,$ belongs to $\Omega$.
+
+\item For each interval $(x_{i-1},x_i), \; 1\leq i \leq m,$ calculate the characteristic
+%\begin{equation}\label{R_i}
+\[
+\begin{gathered}
+R(i)=(1+\gamma_i+\gamma_{i-1} )\Delta_i+(1-\gamma_i-\gamma_{i-1} ) \frac{ (z_i-z_{i-1} )^2}{(r\mu)^2 \Delta_i }-\\
+-\frac{2[(1-\gamma_i )(1+\gamma_{i-1} ) z_i+(1+\gamma_i )(1-\gamma_{i-1} ) z_{i-1} ]}{r\mu},
+\end{gathered}
+%\end{equation}
+\]
+where $r>1$ is the method parameter.
+
+\item Determine the interval to which the maximum characteristic corresponds 
+%\begin{equation}\label{R_max}
+\[
+R(t)=\max\left\{R(i): 1 \leq i \leq m\right\}.
+%\end{equation}
+\]
+
+\item Carry out next trial at the point
+%\begin{equation}\label{new_x}
+\[
+x^{k+1}=\frac {(x_t+x_{t-1})}{2}-(1-\gamma_t-\gamma_{t-1} ) \frac{(z_t-z_{t-1})}{2r\mu}.
+%\end{equation}
+\]
+
+\item Check termination condition $x_t-x_{t-1} \leq \epsilon$, where $t$ is the number of the interval with the maximal characteristic and $\epsilon > 0$ is the predefined accuracy.
+
+\end{enumerate}
+
+Note that the erroneous assignment of some discontinuity points from the series (\ref{points}) will not lead to the loss of convergence of the method.
+The algorithm will interpret such points as a recoverable discontinuity case since the left and right limits of the objective function at such points coincide.
+
+It can be proved that the conditions for the convergence of this algorithm will correspond to the conditions for the convergence of its basic prototype -- the global search algorithm from \cite{Strongin2000}.
+The formulation and rigorous proof of the convergence conditions will be the subject of further publications.
+
+Let us illustrate the operation of the algorithm when searching for the minimum of the function
+\begin{equation}\label{testFunction}
+\varphi(x) =  0.1\sum_{i=1}^5{i\sin{(10(i+1)x+i)}}+
+\begin{cases}
+	-4, &0 \leq x < 3.1\\
+	10, &3.1 \leq x < 4.6\\
+	0, &4.6 \leq x < 7.0\\
+	30, &7.0 \leq x < 9.0\\
+	0, &9.0 \leq x \leq 10.0
+\end{cases}
+\end{equation}
+at $x\in[0,10]$.
+All discontinuity points were considered known, the parameters of the method $r=2.0,\epsilon=0.001$ were used.
+Fig.~\ref{ris1} shows a graph of this function, the discontinuities are marked with a dotted line.
+The dashes under the graph indicate the points of 74 trials required for the algorithm to solve problem with the specified accuracy.
+
+\begin{figure}[!ht]
+	\begin{center}
+			\includegraphics[width=0.9\linewidth]{ris_1.jpg}
+			\caption{Discontinuous objective function graph and trial points in case of given discontinuity points}
+      \label{ris1}
+	\end{center}
+\end{figure}
+
+\section{Identifying Unspecified Discontinuities}
+
+Let us suppose now that the location of the discontinuity points for the function $\varphi(x)$ is not known, but it is known that such points are possible.
+Let the function have a discontinuity at an a priori unknown point $\omega \in(x_{i-1},x_i)$, then the relative difference $\mu_i$ from (\ref{mu_i}) will significantly exceed the relative differences for other intervals since within them the function satisfies the Lipschitz condition.
+Therefore, the identification of such situations can be based on a comparison of the relative differences $\mu_i$ from (\ref{mu_i}) corresponding to different search intervals $(x_{i-1},x_i)$.
+
+Let us enumerate the differences $\mu_i$ from (\ref{mu_i}) in decreasing order (by the index in brackets)
+
+\begin{equation}\label{mu_cond}
+\mu(1) \geq \mu(2) \geq ... \geq \mu(m)
+\end{equation}
+and define the minimum number $p$ satisfying the conditions
+\begin{equation}\label{mu_main_cond}
+ \frac{\mu(p)}{\mu(p+1)} \geq Q, \; 1 \leq p < q(m-2(s+1)),
+\end{equation}
+where $Q>1$ and $0<q<1$ are set numbers.
+
+When at the current $k$-th iteration of the method there is no $p$ that satisfies these conditions, it is interpreted as the absence of unspecified discontinuities.
+If such $p$ exists, this situation corresponds to the presence of undefined discontinuities.
+In this case, zero differences $\mu_i$ from (\ref{mu_i}), corresponding to intervals containing specified discontinuity points, are excluded from consideration by the condition $p<m-2(s+1)$.
+
+Note that the value of the difference $\mu_i$ for the interval containing the extremum point can be close to zero, which will lead to the fulfillment of the first inequality from (\ref{mu_main_cond}).
+To eliminate the effect of small relative differences of intervals containing extreme points, the additional constraint $1 \leq p < q(m-2(s+1))$ is introduced in (\ref{mu_main_cond}) to exclude small values of $\mu_i$ from consideration.
+
+Let us select a subset of numbers
+\begin{equation}\label{set_I}
+I=\left\{ i:1 \leq i \leq m, \; \mu_i=\mu(j), \; 1 \leq j \leq p\right\}
+\end{equation}
+of the intervals $(x_{i-1},x_i)$, which correspond to large relative differences $\mu_i$.
+Large difference values are interpreted as the presence of undefined discontinuities in the intervals $(x_{i-1},x_i), i \in I$.
+Moreover, the set $I$ can change in the process of solving the problem.
+
+\section{Algorithm for unspecified discontinuities}
+
+The algorithm is a modification of the algorithm described in Section 2 for given discontinuities and consists of the following.
+
+The first $s+1$ trials are carried out at arbitrary interior points $x^i \in (\omega_{i-1},\omega_i )$,  $1 \leq i \leq s+1$, where the set $\Omega=\{\omega_0,...,\omega_{s+1}\}$ contains the given discontinuity points, as well as boundary points $\omega_0=a$ and $\omega_{s+1}=b$.
+A point for the next trial $x^{k+1}, \; k \geq s+1$, is selected as follows.
+
+\begin{enumerate}
+\item Sort points of a set $\{x^1,...,x^k\} \cup \Omega$ in ascending order of coordinates 
+\begin{equation}\label{x_cond}
+ a=x_0<x_1<...<x_{m-1}<x_m=b,
+\end{equation}
+where $m=s+k+1$.
+Each point is matched with an attribute $\gamma_i$ from (\ref{gamma}) (the presence or absence of a set discontinuity) and with the value of the function $z_i=\varphi(x_i)$ calculated only for $x_i \not\in \Omega$, i.e.
+at $\gamma_i=0$.
+
+\item Calculate the relative differences $\mu_i, \; 1 \leq i \leq m$, from (\ref{mu_i}) and order them in descending order, i.e. build a series (\ref{mu_cond}).
+
+\item Determine $p$ satisfying the inequalities (\ref{mu_main_cond}), where $0<q<1<Q$ are the parameters of the algorithm.
+Construct a set $I$ from (\ref{set_I}), the elements of which are interpreted as numbers of intervals that can contain a discontinuity.
+Generate interval attributes
+\begin{equation}\label{delta_cond}
+\delta_i=
+\begin{cases}
+	\sign{(z_i - z_{i-1})}, &\text{$i \in I$}, \\
+	0, &\text{$i \not \in I$}.
+\end{cases}
+\end{equation}
+The value $\delta_i=-1$ corresponds to a sharp decrease in the function in the interval $(x_{i-1},x_i)$, $\delta_i=1$ -- to an increase, $\delta_i=0$ -- to the lack of jump.
+Note that in accordance with the rules (\ref{gamma}) and (\ref{delta_cond}) for calculating attributes $\gamma_i$ and $\delta_i$ the inequalities
+$ 0 \leq \gamma_i+\gamma_{i-1}+|\delta_i | \leq 1, \; 1 \leq i \leq m, $
+are met.
+
+\item Determine the maximum value
+\begin{equation}\label{mu_new}
+\mu=\max\left\{\left(1-\gamma_i-\gamma_{i-1}-|\delta_i|\right) \frac {|z_i-z_{i-1}|}{(x_i-x_{i-1})}: 1 \leq i \leq m\right\}.
+\end{equation}
+When $\mu=0$, use the value $\mu=1$.
+
+\item For each interval $(x_{i-1},x_i), \; 1 \leq i \leq m$, calculate the characteristic
+%\begin{equation}\label{R_new}
+\[
+\begin{gathered}
+R(i)=(1+\gamma_i+\gamma_{i-1}+|\delta_i |) \Delta_i+(1-\gamma_i-\gamma_{i-1}-|\delta_i |)\frac{(z_i-z_{i-1} )^2}{(r\mu)^2 \Delta_i } - \\
+- \frac{2[((1-\gamma_i )(1+\gamma_{i-1} )(1-\delta_i )) z_i+((1+\gamma_i )(1-\gamma_{i-1} )(1+\delta_i )) z_{i-1} ]}{r\mu},
+\end{gathered}
+%\end{equation}
+\]
+where $r > 1$ is the method parameter, $\mu$ is from (\ref{mu_new}) and $\Delta_i=x_i-x_{i-1}$.
+
+\item Determine the interval corresponding to the maximum characteristic
+%\begin{equation}\label{R}
+\[
+R(t)=\max\left\{R(i): 1 \leq i \leq m\right\}.
+\]
+%\end{equation}
+
+\item Carry out next trial at the point 
+\begin{equation}\label{new_x_2}
+x^{k+1}=\frac {(x_t+x_{t-1})}{2}-(1-\gamma_t-\gamma_{t-1}- |\delta_t|) \frac{(z_t-z_{t-1})}{2r\mu}.
+\end{equation}
+
+\item Check the termination condition $x_t-x_{t-1} \leq \epsilon$, where $t$ is the number of the interval with the maximal characteristic and $\epsilon > 0$ is the predefined accuracy.
+\end{enumerate}
+
+Fig.~\ref{ris2} shows the results of solving the problem from the previous example (\ref{testFunction}), in which the discontinuity points were considered unspecified.
+Here we used the parameters of the method $r=2.0, \epsilon=0.001, Q=3.0, q=0.3$.
+The dashes under the graph indicate the points of 140 trials required for the algorithm to solve problems with the specified accuracy.
+
+\begin{figure}%[!ht]
+	\begin{center}
+			\includegraphics[width=0.9\linewidth]{ris_2.jpg}
+			\caption{Discontinuous objective function graph and trial points in case of unspecified discontinuity points}
+      \label{ris2}
+	\end{center}
+\end{figure}
+
+\begin{figure}%[!ht]
+	\begin{center}
+			\includegraphics[width=0.9\linewidth]{ris_3.jpg}
+			\caption{Discontinuous objective function graph and trial points in case of partially given discontinuity points}
+      \label{ris3}
+	\end{center}
+\end{figure}
+
+The proposed algorithm also allows partial specification of known discontinuities.
+Fig.~\ref{ris3} shows the results of solving the same problem, for which the discontinuities were considered partially given (the discontinuities at the points 4.6 and 9.0 were considered to be known).
+Here we used the same parameters of the method $r=2.0, \epsilon=0.001, Q=3.0, q=0.3$.
+The dashes under the graph indicate the points of 67 trials required for the algorithm to solve problems with the specified accuracy.
+
+\section{Results of Numerical Experiments}
+
+To demonstrate the efficiency of the global search algorithm for discontinuous functions (GSA-D) described in the previous section, we compare it with Direct Search (DS) \cite{Audet}, Simulated Annealing (SA) \cite{Kirkpatrick}, and Genetic Algorithm (GA) \cite{Goldberg} implemented in MATLAB Global Optimization Toolbox \cite{MatlabOTB}.
+
+As the main comparison criterion, we will use the number of search trials $K$ (i.e., the number of calculations of the objective function) performed by the method before it converges.
+
+The algorithms were compared when solving 1000 problems with a discontinuous objective function, which were constructed as follows. First, we generated continuous functions of the form
+\begin{equation}\label{seriaFunction}
+f(x) =  \sum_{j=0}^4{(j+1)A_j\sin{(5\pi jx+j)}+(j+1)B_j\cos(3\pi jx+j)}, x\in[0,1],
+\end{equation}
+where the values of the coefficients $A_j, B_j$ were chosen randomly and uniformly from the interval $[-6,6]$.
+Then 4 discontinuity points were added to the continuous functions.
+The coordinates of these points $\omega_1, \; \omega_2, \; \omega_3, \; \omega_4$ were set randomly and uniformly from the ranges $[0,0.25), \; [0.25,0.5), \; [0.5,0.75), \; [0.75,1]$ respectively.
+The jump values $\delta_1, \; \delta_2, \; \delta_3, \; \delta_4$ at discontinuities were also randomly selected from the range $[-50,50]$.
+These operations produced a discontinuous function of the form
+\begin{equation}\label{seriaFunction_jump}
+\varphi(x) =  f(x)+
+\begin{cases}
+	\delta_1, &0 \leq x < \omega_1\\
+	\delta_2, &\omega_1 \leq x < \omega_2\\
+	\delta_3, &\omega_2 \leq x < \omega_3 \;, \;\;\; x\in[0,1].\\
+	\delta_4, &\omega_3 \leq x < \omega_4\\
+	\delta_1, &\omega_4 \leq x \leq 1.0
+\end{cases}
+\end{equation}
+
+Since the solution of the test problem assumes that the global optimizer $x^*$ is known, its value was preemptively estimated for each function of the series by iterating through all nodes of a uniform grid, the step of which was chosen to be sufficiently small.
+
+The first experiment was performed when solving a series of 1000 problems with continuous functions of the form (\ref{seriaFunction}).
+Table~\ref{tab1} shows the number of problems solved, the average number of iterations performed by different methods for solving the series, and the number of trials performed by the methods during the search.
+The problem was considered solved if the condition $|x^k-x^* | \leq \delta$ was met for the trial point $x^k$, where $x^*$ is the known solution of the problem and $\delta= 10^{-2}$.
+At the same time, the accuracy used in the termination criteria for all methods was $\epsilon = 10^{-3}$.
+
+In order to exclude the influence of inherent randomness when running algorithms from MatLab, the GA, SA, and DS methods were run 5 times for each problem, and the best result was recorded (in terms of the proximity of the found solution to the true one).
+In this case, the GA and SA methods were run with default parameters, while in the DS method the parameter $x_0$ (the starting point of the search) varied from 0 to 1 with a step of 0.2.
+
+With multiple runs, the methods from MATLAB Global Optimization Toolbox solved almost all the problems in the series, while with a single run of these methods the number of correctly solved problems was almost 2 times less.
+At the same time, the GSA-D method successfully solved all the problems in the series (the reliability parameter $r=2.2$ was used for the method).
+At the same time, the number of trials spent was comparable to the DS method and significantly less than the number of trials of the GA and SA methods.
+
+\begin{table}[ht]
+	\caption{Results of solving a series of problems with continuous functions}
+	\label{tab1}
+	\begin{center}
+		\begin{tabular}{c|c|c|c}
+			\hline
+			Method & Problems solved & Average number of iterations & Average number of trials \\
+			\hline
+%			\hline
+                                GSA-D & 1000  &  52  &  53 \\
+                                \hline
+                                SA    &  1000  &  774  &  777  \\
+			\hline
+			GA    &   999  &  24  &  1290  \\
+			\hline
+                                DS    &  970  & 38  &  71  \\
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
+The next experiment was performed when solving a series of 1000 problems with discontinuous functions of the form (\ref{seriaFunction_jump}).
+GSA-D used the parameters $r=3.7, \; Q=3.0, \; q=0.3$, the other experimental conditions were similar to the previous run.
+The results of solving a series of problems with discontinuous functions are shown in Table~\ref{tab2}.
+
+\begin{table}[ht]
+	\caption{Results of solving a series of problems with discontinuous functions}
+	\label{tab2}
+	\begin{center}
+		\begin{tabular}{c|c|c|c}
+			\hline
+			Method & Problems solved & Average number of iterations & Average number of trials \\
+			\hline
+%			\hline
+                                GSA-D & 1000  & 79  &  80 \\
+                                \hline
+                                SA    &  998  &  765  &  770  \\
+			\hline
+			GA    &   993  &  25  &  1310  \\
+			\hline
+                                DS    &  964  & 38  &  71  \\
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
+The results show that the Global Optimization Toolbox methods solve discontinuous problems somewhat worse than continuous ones, while the GSA-D algorithm proposed in the paper successfully solves the entire series of problems.
+The GA and SA algorithms show high reliability of finding the global optimizer after several runs, but these methods significantly lose to the GSA-D algorithm in terms of the number of objective function calculations.
+The DS method, showing a comparable number of trials, is inferior in the number of correctly solved problems.
+
+Summing up, we note that although the methods from the MATLAB Global Optimization Toolbox are stated as methods that work with discontinuous functions, they do not guarantee convergence to the global minimum.
+The convergence rate (in terms of the number of trials) for the GA and SA methods is significantly lower than that for the DS method.
+However, the DS algorithm, while surpassing the GA and SA methods in this indicator, has the lowest reliability among the considered methods.
+The GSA-D method proposed in the article demonstrated both the guaranteed convergence to the global minimizer for the problems with discontinuous functions and the smaller number of trials required to achieve convergence.
 
 \section{Conclusion}
+
+In this paper, we propose an algorithm for finding the minimum of a multiextremal function with jump discontinuities.
+The proposed GSA-D algorithm, unlike many known approaches, does not use the concept of a gradient and provides a search for a global solution to the problem.
+We carried out computational experiments confirming the convergence of the algorithm for a wide class of discontinuous multiextremal problems.
+The convergence conditions of the proposed algorithm, in general, will correspond to the convergence conditions of its prototype -- the global search algorithm for continuous problems, considered in detail in \cite{Strongin2000}.
+Justification and research of the theoretical properties of the proposed GSA-D algorithm will be the subject of future publications.
 

### 2020-12-16 2021/PСT_GergelKozinov/2021_PCT.tex
fix some misprints

@@
 \end{equation}
-having one continuous and one discrete parameter, the plots of the  function $F(\alpha,y,u)$ for different values of the discrete parameter have form from Fig.~\ref{fig:1} 
+having one continuous and one discrete parameter, the plots of the  function $F(\alpha,y,u)$ for different values of the discrete parameter have form shown in Fig.~\ref{fig:1} and, as one can see, there is no need to calculate the value of the function $F(\alpha,y,u)$ for $u=1$.
+
 \begin{figure}
@@
 \end{figure}
-and, as one can see, there is no need to calculate the value of the function $F(\alpha,y,u)$ for $u=1$.
 
@@
 Unfortunately, attempts to develop parallel variants of AGMIS and AMMIS algorithms using existing parallelization methods have not succeeded.
-For example, data parallelization (splitting the computation domain between available computating elements) results in only one processor processing a subdomain of the search domain  containing the sought globally optimal solution of the problem, with other processors performing redundant calculations.
+For example, data parallelization (splitting the computation domain between available computing elements) results in only one processor processing a subdomain of the search domain  containing the sought globally optimal solution of the problem, with other processors performing redundant calculations.
 In \cite{c24,c32} a new approach to parallelizing calculations for solving global optimization problems is proposed: parallelism of calculations is provided by organizing simultaneous calculation of the values of the minimized function $F(\alpha,y,u)$ from (\ref{eq:8}) at several different points of the search domain $D$.

### 2020-12-16 2021/ICCS_BarkalovUsova/iccs.tex
added actual results

@@
 Fig.~\ref{ris1} shows a graph of this function, the discontinuities are marked with a dotted line.
-The dashes under the graph indicate the points of 74 trials required for the algorithm to solve problem with the specified accuracy.
+The dashes under the graph indicate the points of 70 trials required for the algorithm to solve problem with the specified accuracy.
 
@@
 Here we used the parameters of the method $r=2.0, \epsilon=0.001, Q=3.0, q=0.3$.
-The dashes under the graph indicate the points of 140 trials required for the algorithm to solve problems with the specified accuracy.
+The dashes under the graph indicate the points of 70 trials required for the algorithm to solve problems with the specified accuracy.
 

### 2020-12-24 2021/PСT_GergelKozinov/2021_PCT.tex
1

@@
 
-\institute{Institute of Informational Technologies, Mathematics and Mechanics, Mathematical Center, Lobachevsky State University of Nizhny Novgorod, Nizhni Novgorod, Russia \\
+\institute{Lobachevsky State University of Nizhny Novgorod, Nizhni Novgorod, Russia \\
 \email{gergel@unn.ru, evgeny.kozinov@itmm.unn.ru}
@@
 
-Multicriteria optimization (MCO) problems arise almost every time when the optimal options need to be chosen when developing complex technical devices and systems.
+Multicriteria optimization (MCO) problems arise almost every time when the optimal decision need to be chosen when developing complex technical devices and systems.
 The scale of demand for MCO problems determines the high intensity of research in this area -- see, for example, monographs \cite{c1,c2,c3,c4,c5} and reviews of scientific and practical results in this area \cite{c6,c7,c8,c9}.
@@
 \begin{equation}\label{eq:10}
-\mathbb{F}_T=\{F(\alpha_t,y):1 \leq t \leq T\},
+\mathbb{F}_T=\{F(\alpha_t,y , u):1 \leq t \leq T\},
 \end{equation}
@@
 \textit{Rule 3.} Check the stopping condition for the computation.
-If the required accuracy of the global search is not achieved, it is necessary to supplement the search information $\omega_k$ from (\ref{eq:13}) with the results of the current iteration and continue calculations from Rule 1.
+If the required accuracy of the global search is not achieved, it is necessary to supplement the search information $\omega_k$ from (\ref{eq:13}) with the results of the current iteration and continue calculations starting from Rule 1.
 
@@
 ($E(x)$ denotes the operation of taking the integral part of the number $x$).
-It should be noted that the function $\Phi(\alpha,y(x))$ is discontinuous at the points $x_i$, $1 \leq i \leq l-1$,  further on, the values of the function $\Phi(\alpha,y(x))$ at these points are considered undefined and not used in the calculations.
+It should be noted that the function $\Phi(\alpha,y(x))$ is discontinuous at the points $x_i$, $1 \leq i \leq l-1$,  further on, the values of the function $\Phi(\alpha,y(x))$ at these points are considered undefined and are not used in the calculations.
 
@@
 Next, let $k$, $k > 1$, global search iterations be performed.
-Selection of the trial points $(k+1)$ for the next iteration is determined by the following rules.
+Selection of the trial point $(k+1)$ for the next iteration is determined by the following rules.
 
@@
 To solve each MCOmix problem, 50 uniformly distributed values of convolution coefficients $\lambda \in \Lambda$ from (\ref{eq:8}) were used.
-To estimate the values of HV and DU efficiency indicators, an approximation of the Pareto domain was calculated for each MCOmix problem  using a uniform grid in the domain of varied values $D\times U$ from (\ref{eq:2})-(\ref{eq:3}).
+To estimate the values of HV and DU efficiency indicators, an approximation of the Pareto domain was calculated for each MCOmix problem  using a uniform grid in the search domain $D\times U$ from (\ref{eq:2})-(\ref{eq:3}).
 The PAMMIS algorithm was used for the following parameter values: accuracy $\varepsilon=0.05$ from (\ref{eq:26}) and reliability $r=5.6$ (the $r$ parameter is used in PAMMIS to construct estimates of the H{\"o}lder constant $H$ from (\ref{eq:16})).
@@
 \multicolumn{7}{c}{\begin{tabular}[c]{@{}c@{}}The Pareto domain estimate obtained \\ by exhaustive search\end{tabular}}        \\ \hline
-48                     & 1 706 667                   & -                       & 20,6         & &       & 29 784,9                \\ \hline
+48                     & 1706667                   & -                       & 20.6         & &       & 29784.9                \\ \hline
 \multicolumn{7}{c}{\begin{tabular}[c]{@{}c@{}}The Pareto    domain  estimate obtained \\ using the PAMMIS method\end{tabular}} \\ \hline
-1                      & 86 261,6                    & 1,0                     & 13,8         & &      & 30 212,3                \\
-6                      & 15 406,2                    & 5,6                     & 14,4         & &      & 30 317,6                \\
-12                     & 6 726,1                     & 12,8                    & 16,2         & &      & 30 248,3                \\
-18                     & 5 972,7                     & 14,4                    & 14,3         & &      & 30 551,1                \\
-24                     & 3 657,4                     & 23,6                    & 13,9         & &      & 30 092,3                \\
-30                     & 3 162,3                     & 27,3                    & 14,9         & &      & 30 443,9                \\
-36                     & 2 888,7                     & 29,9                    & 14,6         & &      & 30 528,7                \\
-42                     & 2 028,8                     & 42,5                    & 15,6         & &      & 30 046,1                \\
-48                     & 1 767,6                     & 48,8                    & 15,3         & &      & 30 255,5                \\ \hline
+1                      & 86261.6                    & 1.0                     & 13.8         & &      & 30212.3                \\
+6                      & 15406.2                    & 5.6                     & 14.4         & &      & 30317.6                \\
+12                     & 6726.1                     & 12.8                    & 16.2         & &      & 30248.3                \\
+18                     & 5972.7                     & 14.4                    & 14.3         & &      & 30551.1                \\
+24                     & 3657.4                     & 23.6                    & 13.9         & &      & 30092.3                \\
+30                     & 3162.3                     & 27.3                    & 14.9         & &      & 30443.9                \\
+36                     & 2888.7                     & 29.9                    & 14.6         & &      & 30528.7                \\
+42                     & 2028.8                     & 42.5                    & 15.6         & &      & 30046.1                \\
+48                     & 1767.6                     & 48.8                    & 15.3         & &      & 30255.5                \\ \hline
 \end{tabular}
@@
 The results of the computational experiments show that the PAMMIS algorithm is scalable as the  speedup of parallel computations increases almost linearly with the number of computing cores used.
-The obtained values of HV and DU indicators show that the Pareto set estimate using the PAMMIS algorithm is calculated with a greater accuracy and requires a significantly smaller number of trials compared to the results obtained using uniform grids in the domain $D \times U$.
+The obtained values of HV and DU indicators show that the Pareto set estimate using the PAMMIS algorithm is calculated with a greater accuracy and requires a significantly smaller number of trials compared to the results obtained using uniform grids in the search domain $D \times U$.
 
@@
 \section*{Acknowledgements} 
-This work was supported by the project no. 0729-2020-0055 and by the Research and Education Mathematical Center, project no. 075-02-2020-1483/1
+This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.0729-2020-0055, and by the Research and Education Mathematical Center, project no.
+075-02-2020-1483/1.
 
-\bibliographystyle{spmpsci}
-%\bibliographystyle{unsrt}
-\bibliography{mybibfile}
+\begin{thebibliography}{10}
+\providecommand{\url}[1]{{#1}}
+\providecommand{\urlprefix}{URL }
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
+  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
+  \urlstyle{rm}\Url}\fi
+
+\bibitem{c45}
+Barkalov, K., Lebedev, I.: Parallel global optimization for non-convex
+  mixed-integer problems.
+\newblock Communications in Computer and Information Science \textbf{1129},
+  98--109 (2019).
+\newblock \doi{10.1007/978-3-030-36592-9_9}
+
+\bibitem{c15}
+Belotti, P., Lee, J., Liberti, L., Margot, F., W{\" a}chter, A.: Branching and
+  bounds tightening techniques for non-convex {MINLP}.
+\newblock Optimization Methods and Software \textbf{24}(4-5), 597--634 (2009).
+\newblock \doi{10.1080/10556780903087124}
+
+\bibitem{c14}
+Boukouvala, F., Misener, R., Floudas, C.: {Global optimization advances in
+  mixed-integer nonlinear programming, MINLP, and constrained derivative-free
+  optimization, CDFO}.
+\newblock European Journal of Operational Research \textbf{252}(3), 701--727
+  (2016).
+\newblock \doi{10.1016/j.ejor.2015.12.01}
+
+\bibitem{c10}
+Branke, J., Deb, K., Miettinen, K., Slowinski, R.: Multiobjective optimization:
+  interactive and evolutionary approaches, vol. 5252.
+\newblock Springer, Berlin (2008).
+\newblock \doi{10.1007/978-3-540-88908-3}
+
+\bibitem{c13}
+Burer, S., Letchford, A.: Non-convex mixed-integer nonlinear programming: a
+  survey.
+\newblock Surveys in Operations Research and Management Science \textbf{17},
+  97--106 (2012).
+\newblock \doi{10.1016/j.sorms.2012.08.001}
+
+\bibitem{c3}
+Collette, Y., Siarry, P.: Multiobjective optimization: principles and case
+  studies (decision engineering).
+\newblock Springer-Verlag, Berlin, Heidelberg (2011)
+
+\bibitem{c17}
+Deep, K., Singh, K., Kansal, M., Mohan, C.: A real coded genetic algorithm for
+  solving integer and mixed integer optimization problems.
+\newblock Applied Mathematics and Computation \textbf{212}(2), 505--518 (2009).
+\newblock \doi{10.1016/j.amc.2009.02.044}
+
+\bibitem{c2}
+Ehrgott, M.: Multicriteria optimization.
+\newblock Springer-Verlag, Berlin, Heidelberg (2005)
+
+\bibitem{c12}
+Eichfelder, G.: Scalarizations for adaptively solving multi-objective
+  optimization problems.
+\newblock Coldomput. Optim. Appl. \textbf{44}, 249--273 (2009).
+\newblock \doi{10.1007/s10589-007-9155-4}
+
+\bibitem{c42}
+Evtushenko, Y., Posypkin, M.: Method of non-uniform coverages to solve the
+  multicriteria optimization problems with guaranteed accuracy.
+\newblock Automation and Remote Control \textbf{75}(6), 1025--1040 (2014).
+\newblock \doi{10.1134/S0005117914060046}
+
+\bibitem{c31}
+Floudas, C., Pardalos, M.: Recent advances in global optimization.
+\newblock Princeton University Press (2016)
+
+\bibitem{c44}
+Gaviano, M., Kvasov, D., Lera, D., Sergeyev, Y.: Software for generation of
+  classes of test functions with known local and global minima for global
+  optimization.
+\newblock ACM Transactions on Mathematical Software \textbf{29}(4), 469--480
+  (2003)
+
+\bibitem{c19}
+Gergel, V.: A unified approach to use of coprocessors of various types for
+  solving global optimization problems.
+\newblock 2nd International Conference on Mathematics and Computers in Sciences
+  and in Industry pp. 13--18 (2015).
+\newblock \doi{10.1109/MCSI.2015.18}
+
+\bibitem{c23}
+Gergel, V., Barkalov, K., Lebedev, I.: A global optimization algorithm for
+  non-convex mixed-integer problems.
+\newblock Lecture Notes in Computer Science \textbf{11353}, 78--81 (2019).
+\newblock \doi{10.1007/978-3-030-05348-2_7}
+
+\bibitem{c41}
+Gergel, V., Barkalov, K., Sysoyev, A.: A novel supercomputer software system
+  for solving time-consuming global optimization problems.
+\newblock Numerical Algebra, Control \& Optimization \textbf{8}(1), 47--62
+  (2018)
+
+\bibitem{c20}
+Gergel, V., Kozinov, E.: Accelerating parallel multicriterial optimization
+  methods based on intensive using of search information.
+\newblock Procedia Computer Science \textbf{108}, 1463--1472 (2017).
+\newblock \doi{10.1016/j.procs.2017.05.051}
+
+\bibitem{c21}
+Gergel, V., Kozinov, E.: Efficient multicriterial optimization based on
+  intensive reuse of search information.
+\newblock J. Glob. Optim. \textbf{71}(1), 73--90 (2018).
+\newblock \doi{10.1007/s10898-018-0624-3}
+
+\bibitem{c22}
+Gergel, V., Kozinov, E.: Multilevel parallel computations for solving
+  multistage multicriteria optimization problems.
+\newblock Lecture Notes in Computer Science \textbf{12137}, 17--30 (2020)
+
+\bibitem{c7}
+Greco, S., Ehrgott, M., Figueira, J.: Multiple criteria decision analysis:
+  state of the art surveys.
+\newblock Springer (2016).
+\newblock \doi{10.1007/978-1-4939-3094-4}
+
+\bibitem{c38}
+Grishagin, V., Israfilov, R., Sergeyev, Y.: Comparative efficiency of
+  dimensionality reduction schemes in global optimization.
+\newblock AIP Conference Proceedings \textbf{1776}, 060011 (2016).
+\newblock \doi{10.1063/1.4965345}
+
+\bibitem{c40}
+Grishagin, V., Sergeyev, Y., Strongin, R.: Parallel characteristical global
+  optimization algorithms.
+\newblock Journal of Global Optimization \textbf{10}, 185--206 (1997)
+
+\bibitem{c9}
+Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and
+  industrial applications.
+\newblock Surveys on Mathematics for Industry \textbf{11}, 1--42 (2005)
+
+\bibitem{c36}
+Lera, D., Sergeyev, Y.: Lipschitz and {H}{\"o}lder global optimization using
+  space-filling curves.
+\newblock Applied Numerical Mathematics \textbf{60}(1-2), 115--129 (2010).
+\newblock \doi{10.1016/j.apnum.2009.10.004}
+
+\bibitem{c29}
+Locatelli, M., Schoen, F.: Global optimization: theory, algorithms, and
+  applications.
+\newblock {SIAM} (2013)
+
+\bibitem{c6}
+Marler, R., Arora, J.: Survey of multi-objective optimization methods for
+  engineering.
+\newblock Structural and Multidisciplinary Optimization \textbf{26}, 369--395
+  (2004).
+\newblock \doi{10.1007/s00158-003-0368-6}
+
+\bibitem{c4}
+Marler, R., Arora, J.: Multi-objective optimization: concepts and methods for
+  engineering (2009)
+
+\bibitem{c1}
+Miettinen, K.: Nonlinear multiobjective optimization.
+\newblock Springer (1998)
+
+\bibitem{c5}
+Pardalos, P., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-convex
+  multi-objective optimization, vol. 123.
+\newblock Springer (2017)
+
+\bibitem{c30}
+Paulavi{\v c}ius, R., {\v Z}ilinskas, J.: Simplicial global optimization.
+\newblock Springer-Verlag, New York (2014).
+\newblock \doi{10.1007/978-1-4614-9093-7}
+
+\bibitem{c27}
+Pint\'er, J.: Global optimization in action (continuous and {Lipschitz}
+  optimization: algorithms, implementations and applications).
+\newblock Kluwer Academic Publishers, Dordrecht (1996)
+
+\bibitem{c18}
+Schl\"{u}ter, M., Egea, J., Banga, J.: Extended ant colony optimization for
+  non-convex mixed integer nonlinear programming.
+\newblock Comput. Oper. Res. \textbf{36}(7), 2217–2229 (2009).
+\newblock \doi{10.1016/j.cor.2008.08.015}
+
+\bibitem{c33}
+Sergeyev, Y.: An information global optimization algorithm with local tuning.
+\newblock SIAM Journal on Optimization \textbf{5}(4), 858--870 (1995).
+\newblock \doi{10.1137/0805041}
+
+\bibitem{c35}
+Sergeyev, Y., Famularo, D., Pugliese, P.: Index branch-and-bound algorithm for
+  global optimization with multiextremal constraints.
+\newblock Journal of Global Optimization \textbf{21}(3), 317--341 (2001).
+\newblock \doi{10.1023/A:1012391611462}
+
+\bibitem{c34}
+Sergeyev, Y., Grishagin, V.: Parallel asynchronous global search and the nested
+  optimization scheme.
+\newblock Journal of Computational Analysis and Applications \textbf{3}(2),
+  123--145 (2001).
+\newblock \doi{10.1023/A:1010185125012}
+
+\bibitem{c37}
+Sergeyev, Y., Kvasov, D.: A deterministic global optimization using smooth
+  diagonal auxiliary functions.
+\newblock Communications in Nonlinear Science and Numerical Simulation
+  \textbf{21}(1-3), 99--111 (2015).
+\newblock \doi{10.1016/j.cnsns.2014.08.026}
+
+\bibitem{c39}
+Sergeyev, Y., Nasso, M., Mukhametzhanov, M., Kvasov, D.: Novel local tuning
+  techniques for speeding up one-dimensional algorithms in expensive global
+  optimization using {Lipschitz} derivatives.
+\newblock Journal of Computational and Applied Mathematics \textbf{383} (2021).
+\newblock \doi{10.1016/j.cam.2020.113134}
+
+\bibitem{c25}
+Sergeyev, Y., Strongin, R., Lera, D.: Introduction to global optimization
+  exploiting space-filling curves.
+\newblock Springer, New York (2013)
+
+\bibitem{c32}
+Strongin, R., Sergeyev, Y.: Global multidimensional optimization on parallel
+  computer.
+\newblock Parallel Computing \textbf{18}(11), 1259--1273 (1992).
+\newblock \doi{10.1016/0167-8191(92)90069-J}
+
+\bibitem{c24}
+Strongin, R., Sergeyev, Y.: Global optimization with non-convex constraints.
+  Sequential and parallel algorithms.
+\newblock Kluwer Academic Publishers, Dordrecht (2000 (2nd ed. 2013, 3rd ed.
+  2014))
+
+\bibitem{c16}
+Vigerske, S., Gleixner, A.: {SCIP}: global optimization of mixed-integer
+  nonlinear programs in a branch-and-cut framework.
+\newblock Optimization Methods and Software \textbf{33}(3), 563--593 (2018).
+\newblock \doi{10.1080/10556788.2017.1335312}
+
+\bibitem{c11}
+Voutchkov, I., Keane, A.: Multi-objective optimization using surrogates.
+\newblock Computational Intelligence in Optimization \textbf{7}, 155--175
+  (2010).
+\newblock \doi{10.1007/978-3-642-12775-5_7}
+
+\bibitem{c8}
+Zavadskas, E., Turskis, Z., Kildiene, S.: State of art surveys of overviews on
+  {MCDM}/{MADM} methods.
+\newblock Technological and Economic Development of Economy \textbf{20},
+  165--179 (2014).
+\newblock \doi{10.3846/20294913.2014.892037}
+
+\bibitem{c26}
+Zhigljavsky, A.: Theory of global random search.
+\newblock Kluwer Academic Publishers, Dordrecht (1991)
+
+\bibitem{c28}
+Zhigljavsky, A., {\v Z}ilinskas, A.: Stochastic global optimization, vol.~9.
+\newblock Springer, Berlin (2008).
+\newblock \doi{10.1007/978-0-387-74740-8}
+
+\bibitem{c43}
+{\v Z}ilinskas, A., {\v Z}ilinskas, J.: Adaptation of a one-step worst-case
+  optimal univariate algorithm of bi-objective {Lipschitz} optimization to
+  multidimensional problems.
+\newblock Commun Nonlinear Sci Numer Simulat \textbf{21}(1-3), 89--98 (2015).
+\newblock \doi{10.1016/j.cnsns.2014.08.025}
+
+\end{thebibliography}
 

### 2021-01-19 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
MOTOR2021 paper template

--- 2018-12-28 2019/MOTOR/motor.tex
@@
 
+%Пакеты для отображения русского шрифта
+\usepackage[utf8]{inputenc}
+\usepackage[russian]{babel}
+%Пакеты для гиперссылок
+\usepackage{url}
+\usepackage{hyperref}
+\usepackage{cite}
+
 \begin{document}
 %
-\title{ Generalized nested optimization scheme for global optimization problems
-\thanks{This study was supported by the Russian Science Foundation, project 
-No.\,16-11-10150.}}
+\title{Global search algorithm with local tuning for solving mixed-integer global optimization problems
+%with non-convex constraints 
+\thanks{This study was supported by the .}}
 %
@@
 %
-\author{Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
-Ilia Lebedev\orcidID{0000-0002-8736-0652}}
+\author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
+Ilya Lebedev\orcidID{0000-0002-8736-0652}
+}
 %
@@
 %
-\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
-\email{konstantin.barkalov@itmm.unn.ru}}
+\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
+\email{\{konstantin.barkalov,ilya.lebedev\}@itmm.unn.ru}}
 %
@@
 \subsection{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is not indented. The first paragraph that follows a table, figure,
+Что-то по русски
+Please note that the first paragraph of a section or subsection is
+not indented. The first paragraph that follows a table, figure,
 equation etc. does not need an indent, either.
@@
 \end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
+Please try to avoid rasterized images for line-art diagrams and schemas. Whenever possible, use vector graphics instead (see Fig.~\ref{fig1}).
 
@@
 \end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{globalizerSystem}, a
-book~\cite{strSergGO}, proceedings without editors~\cite{Jones2009}.
+\cite{Strongin2000}.
 %
@@
 %
-
-%\begin{thebibliography}{8}
-%\bibitem{ref_article1}
-%Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-%
-%\bibitem{ref_lncs1}
-%Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-%F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-%Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-%
-%\bibitem{ref_book1}
-%Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-%Location (1999)
-%
-%\bibitem{ref_proc1}
-%Author, A.-B.: Contribution title. In: 9th International Proceedings
-%on Proceedings, pp. 1--2. Publisher, Location (2010)
-%
-%\bibitem{ref_url1}
-%LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-%Oct 2017
-%\end{thebibliography}
 \end{document}

### 2021-01-26 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
template for LJM paper

%%
%% ****** ljmsamp.tex 13.06.2018 ******
%%
\documentclass[
11pt,%
tightenlines,%
twoside,%
onecolumn,%
nofloats,%
nobibnotes,%
nofootinbib,%
superscriptaddress,%
noshowpacs,%
centertags]%
{revtex4}
\usepackage{ljm}

\begin{document}

\titlerunning{Short form of the title} % for running heads
%\authorrunning{First-Author at al.} % for running heads
\authorrunning{Barkalov, Lebedev} % for running heads

\title{Title of the Article,\\
Broken into Lines}
% Splitting into lines is performed by the command \\
% The title is written in accordance with the rules of capitalization.

\author{\firstname{K.~A.}~\surname{Barkalov}}
\email[E-mail: ]{konstantin.barkalov@itmm.unn.ru}
\affiliation{Lobachevsky State University of Nizhni Novgorod, Gagarin ave. 23, Nizhni Novgorod, 603950 Russia}

\author{\firstname{I.~G.}~\surname{Lebedev}}
\email[E-mail: ]{ilya.lebedev@itmm.unn.ru}
\affiliation{Lobachevsky State University of Nizhni Novgorod, Gagarin ave. 23, Nizhni Novgorod, 603950 Russia}
%\noaffiliation % If the author does not specify a place of work.

\firstcollaboration{(Submitted by A.~A.~Editor-name)} % Add if you know submitter.
%\lastcollaboration{ }

\received{February 1, 2021} % The date of receipt to the editor, i.e. December 06, 2017

\begin{abstract} % You shouldn't use formulas and citations in the abstract.
In this example, the article contains some required author
information and examples of how to gain an article in the REV\TeX~4
for \ljm. You shouldn't use formulas and citations in the abstract.
\end{abstract}

\subclass{90C26, 90C30} % Enter 2010 Mathematics Subject Classification.

\keywords{Global optimization, Non-convex constraints, Mixed-integer problems, Local tuning, Parallel algorithms} % Include keywords separeted by comma.

\maketitle

% Text of article starts here.

\section{Introduction}

To prepare your manuscript, use REV\TeX{} package. The latest version can be downloaded at the project homepage \cite{RTeXHome}.
The process of processing an article using REV\TeX{} is described in detail in the package manuals \cite{RTeX}. Great help in resolving technical
issues on \TeX{} can be found in books \cite{texbook,L,GG,KD}.
Use the ``House Style Guide: Version 2.0'' \cite{hsg} for more help on how to format article.

\section{First level Heading\protect\\
broken into lines}

There are three levels of Headings that are set by the commands
\verb|\section|, \verb|\subsection|, and \verb|\subsubsection|
(Chapter, subchapter, subsection). Split header into lines
is done with the command \verb+\protect\\+.
Second level heading as the title of article is written in accordance with the rules of capitalization \cite{cap}.

\subsection{Second Level Heading}

References in the text are by using the commands
\verb+\cite{#1}+ or \verb+\onlinecite{#1}+. Label \verb+#1+
can have a name consisting of both letters and numbers. In
the bibliography section this link also has a label \verb+#1+ and starts the command \verb+\bibitem{#1}+.

With the command \verb+\cite{texbook}+ get a reference \cite{texbook},
if you need to refer to several sources:
\cite{texbook,L}, the curly brackets indicate references
separated by commas: \verb+\cite{texbook,L}+. When
use the command \verb|\onlinecite{#1}| link will not be
in brackets: see~\onlinecite{texbook}, p.~8.
If there is a link to the sources with serial numbers,
e.g. [1,2,3,6,7,8], then print automatically following links will take
[1--3,6--8].

\section{Standalone formulae}

\subsection{Another second-level heading}

In \LaTeX, there are many ways to embed standalone formulas
on the page and align them. By default, formulas always
centered.

\subsubsection{One-line formulas}

Below are examples of one-line equations:
\begin{eqnarray}
\chi_+(p)\alt{\bf [}2|{\bf p}|(|{\bf p}|+p_z){\bf ]}^{-1/2}
\left(
\begin{array}{c}
|{\bf p}|+p_z\\
px+ip_y
\end{array}\right)\;,
\\
\left\{%
\openone234567890abc123\alpha\beta\gamma\delta1234556\alpha\beta
\frac{1\sum^{a}_{b}}{A^2}%
\right\}%
\label{one}.
\end{eqnarray}
The second formula has the number~(\ref{one}), which is set by command
\verb|\label{one}|. The first formula is assigned a number~(1), but
it cannot be applied with the help of automatic links,
as it has no label.

If the formula number is not necessary, use the environment
\verb+\[+, \verb+\]+, (or \$...\$) which obtained the following formula:
\[g^+g^+ \rightarrow g^+g^+g^+g^+ \dots ~,~~q^+q^+\rightarrow
q^+g^+g^+ \dots ~. \] % You also can use the dollar sign ($).

\subsubsection{Multiline formulae}

Multiline formulas are typed using the environment
eqnarray:
\begin{eqnarray}
{\cal M}=&&ig_Z^2(4E_1E_2)^{1/2}(l_i^2)^{-1}
\delta_{\sigma_1,-\sigma_2}
(g_{\sigma_2}^e)^2\chi_{-\sigma_2}(p_2)\nonumber\\
&&\times
[\epsilon_jl_i\epsilon_i]_{\sigma_1}\chi_{\sigma_1}(p_1),
\end{eqnarray}
\begin{eqnarray}
\sum \vert M^{\rm viol}_g \vert ^2&=&g^{2n-4}_S(Q^2)~N^{n-2}
(N^2-1)\nonumber \\
 & &\times \left( \sum_{i<j}\right)
\sum_{\rm perm}
\frac{1}{S_{12}}
\frac{1}{S_{12}}
\sum_\tau c^f_\tau~.
\end{eqnarray}
If the formula number is not necessary, then at the end of the row in front of the sign
\verb|\\| you need to put the command \verb|\nonumber|. Never
use one line command \verb|\nonumber| and
\verb|\label{#1}|, as this may cause error in automatic
the numbering of references.

If you want to gain a few formulas without number,
use the eqnarray environment* (the asterisk means the abolition
numbering):
\begin{eqnarray*}
\sum \vert M^{\rm viol}_g \vert ^2&=&g^{2n-4}_S(Q^2)~N^{n-2}
(N^2-1)\\
& &\times \left( \sum_{i<j}\right)
\left(
\sum_{\rm perm}\frac{1}{S_{12}S_{23}S_{n1}}
\right)
\frac{1}{S_{12}}~.
\end{eqnarray*}
To add numbers to the formula manually, use the command
\verb+\tag{#1}+, where \verb+#1+~--- the desired equation number.
Here how is the formula with the number of~(\ref{eq:mynum}):
\begin{equation}
g^+g^+ \rightarrow g^+g^+g^+g^+ \dots ~,~~q^+q^+\rightarrow
q^+g^+g^+ \dots ~. \tag{2.6$'$}\label{eq:mynum}
\end{equation}

When you enable single-line and multi-line formulas are surrounded by
subequations, each formula `numbered" with a letter,
as shown in equations~(\ref{mlett:1}) and (\ref{mlett:2}):
\begin{subequations}
\label{generallabel}
\begin{equation}
\left\{
abc123456abcdef\alpha\beta\gamma\delta1234556\alpha\beta
\frac{1\sum^{a}_{b}}{A^2}
\right\},\label{mlett:1}
\end{equation}
\begin{eqnarray}
{\cal M}=&&ig_Z^2(4E_1E_2)^{1/2}(l_i^2)^{-1}
(g_{\sigma_2}^e)^2\chi_{-\sigma_2}(p_2)\nonumber\\
&&\times
[\epsilon_i]_{\sigma_1}\chi_{\sigma_1}(p_1).\label{mlett:2}
\end{eqnarray}
\end{subequations}
If you put the label right after the \verb+\begin{subequations}+,
it can be used further as a reference to all the equations in this
environment. For example, you can refer to
equation~(\ref{generallabel}) this example.

To set multi-line formulas you can use the environment
multline, gather and align. The multline environment is good to use
for a set of standalone long formulas that do not fit on
one line:
\begin{multline}
\int_{a_1}^{a_2} f(x)\,dx+\int_{a_2}^{a_3} f(x)\,dx
+\dots+\int_{a_{n-1}}^{a_n} f(x)\,dx\\
+\int_{a_1}^{a_2} g(x)\,dx+\int_{a_2}^{a_3} g(x)\,dx
+\dots+\int_{a_{n-1}}^{a_n} g(x)\,dx\\
+\int_{a_1}^{a_2} h(x)\,dx+\int_{a_2}^{a_3} h(x)\,dx
+\dots+\int_{a_{n-1}}^{a_n} h(x)\,dx\\
=\int_{a_1}^{a_n} f(x)+g(x)+h(x)\,dx.
\end{multline}
This formula is automatically numbered, if the formula number is not
need, you have to use the environment multline*.

Environment gather centers included in the formula:
\begin{gather}
\int_{a_1}^{a_2} f(x)\,dx+\int_{a_2}^{a_3} f(x)\,dx
+\dots+\int_{a_{n-1}}^{a_n} f(x)\,dx\\
+\int_{a_1}^{a_2} g(x)\,dx+\int_{a_2}^{a_3} g(x)\,dx
+\dots+\int_{a_{n-1}}^{a_n} g(x)\,dx\notag\\
+\int_{a_1}^{a_2} h(x)\,dx+\int_{a_2}^{a_3} h(x)\,dx
+\dots+\int_{a_{n-1}}^{a_n} h(x)\,dx\\
=\int_{a_1}^{a_n} f(x)+g(x)+h(x)\,dx.
\end{gather}
Each line is automatically numbered, if the line number
it is not necessary, before \verb+\\+ in this line, you need to put the command
\verb+\notag+. When you use the environment gather* formula
will not be numbered.

The align environment allows you to align formulas on your
discretion:
\begin{align}
\int_{a_1}^{a_2} f(x)\,dx &+\int_{a_2}^{a_3} f(x)\,dx
+\dots+\int_{a_{n-1}}^{a_n} f(x)\,dx\notag\\
&+\int_{a_1}^{a_2} g(x)\,dx +\int_{a_2}^{a_3} g(x)\,dx
+\dots+\int_{a_{n-1}}^{a_n} g(x)\,dx\\
&+\int_{a_1}^{a_2} h(x)\,dx+\int_{a_2}^{a_3} h(x)\,dx
+\dots+\int_{a_{n-1}}^{a_n} h(x)\,dx\notag\\
&=\int_{a_1}^{a_n} f(x)+g(x)+h(x)\,dx.
\end{align}
Read more about working with these environments can be found in the book
\cite{GG}.

\section{Variables}
For the writing of definitions, theorems, lemmas and their proofs use the following variables. If necessary, you can add your variables in the sample.

\begin{definition}\label{D:1}
Define ...
\end{definition}

\begin{lemma}\label{L:1}
If ...
\end{lemma}

\begin{theorem}\label{Th:1}
Let ...
\end{theorem}
\begin{proof}
Consider...
\end{proof}

\section{Figures and tables}
REV\TeX~4 will automatically number sections, equations, tables, and
drawings. Figure captions are made after the image and the signature
tables~--- before a table, as shown in the examples at the end. In
tables footnotes \footnote{Footnotes in tables, you can try
to do this manually.} not working.

Graphics and diagrams should be monochrome in black-and-white color and saved in EPS format.
%You can include a picture in the article using the figure environment:
%\begin{figure}[h]
%\setcaptionmargin{5mm}
%%\onelinecaptionsfalse % if the caption is multiline
%\onelinecaptionstrue  % if the caption is one-line
%\includegraphics[width=0.85\textwidth]{deform.eps}
%\captionstyle{normal}\caption{Please write your figure caption here.}\label{fig:1}
%\end{figure}

\begin{table}[!h]
\setcaptionmargin{0mm}
\onelinecaptionsfalse
\captionstyle{flushleft}
\caption{ For the insertion of tables, the table environment is used,
 the signatures to the tables are made in the same way as the captions
 for the figures. This is an example of a table whose multi-line name
 is decorated with the \textbf{caption2} package.}
\bigskip
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
 &$r_c$ (\AA)&$r_0$ (\AA)&$\kappa r_0$&
 &$r_c$ (\AA) &$r_0$ (\AA)&$\kappa r_0$\\
\hline
Cu& 0.800 & 14.10 & 2.550 &Sn
& 0.680 & 1.870 & 3.700 \\
Ag& 0.990 & 15.90 & 2.710 &Pb
& 0.450 & 1.930 & 3.760 \\
Au& 1.150 & 15.90 & 2.710 &Ca
& 0.750 & 2.170 & 3.560 \\
Mg& 0.490 & 17.60 & 3.200 &Sr
& 0.900 & 2.370 & 3.720 \\
Zn& 0.300 & 15.20 & 2.970 &Li
& 0.380 & 1.730 & 2.830 \\
Cd& 0.530 & 17.10 & 3.160 &Na
& 0.760 & 2.110 & 3.120 \\
Hg& 0.550 & 17.80 & 3.220 &K &  1.120 & 2.620 & 3.480 \\
Al& 0.230 & 15.80 & 3.240 &Rb & 1.330 & 2.800 & 3.590 \\
Ga& 0.310 & 16.70 & 3.330 &Cs & 1.420 & 3.030 & 3.740 \\
In& 0.460 & 18.40 & 3.500 &Ba & 0.960 & 2.460 & 3.780 \\
Tl& 0.480 & 18.90 & 3.550 & & & & \\[1mm]
\hline
\end{tabular}
\end{table}

\begin{table}[!htb]
\setcaptionmargin{0mm}
\onelinecaptionstrue
\captionstyle{flushleft}
\caption{The name of this table is --- one-line.}
\bigskip
\begin{tabular}{|c|c|c|c|c|c|c|}
  \hline
    & 1 & 2 & 3 & 4 & 5 & 6\\
  \hline
  1 & 1 & 2 & 3 & 4 & 5 & 6\\
  2 & 2 & 4 & 6 & 8 & 10 & 12\\
  3 & 3 & 6 & 9 & 12 & 15 & 18\\[1mm]
  \hline
\end{tabular}
\end{table}

\section{Bibliography}
The following rules apply for references to books:

\begin{itemize}
  \item Authors initials go before the surname with a space between the initials. Use and between the last two authors. All authors listed in the original reference should be cited.
  \item The title of the book is written in italics. If the book is originally published in Russian, only the English translation of the title is cited and after citation the original language may be indicated in square brackets, e.g.: [in Russian].
  \item Examples of bibliographic references can be found at the end of this page % (i.e. article \cite{ex}, PhD thesis \cite{maguire-76}).
\end{itemize}

\begin{acknowledgments}
%We are grateful to C.~C.~Surname3 and the reviewers for careful reading of the manuscript and helpful remarks.
This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no. 0729-2020-0055, and by the Research and Education Mathematical Center, project no. 075-02-2020-1483/1.
\end{acknowledgments}

%
% The Bibliography
%

\begin{thebibliography}{99}

%\bibitem{Floudas}
%C.~A.~Floudas and M.~P.~Pardalos, \textit{Recent advances in global optimization} (Princeton University Press, 2016).
%
%\bibitem{Locatelli}
%M.~Locatelli and F.~Schoen, \textit{Global optimization: theory, algorithms and applications} (SIAM, 2013).
%
%\bibitem{Strongin1}
%R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global optimization with non-convex constraints. Sequential and parallel algorithms} (Kluwer Academic Publishers, Dordrecht, 2000, 2nd ed. 2013, 3rd ed. 2014).
%
%\bibitem{Pardalos}
%P.~M.~Pardalos, A.~A.~Zhigljavsky and J.~\v{Z}ilinskas \textit{Advances in stochastic and deterministic global optimization} (Springer, 2016).
%
%\bibitem{Sergeyev1}
%Y.~D.~Sergeyev and D.~E.~Kvasov, \textit{Deterministic Global Optimization. An Introduction to the Diagonal Approach}  (Springer Briefs in Optimization, Springer, 2017).
%
%\bibitem{Paulavicius}
%R.~Paulavi\v{c}ius, J.~\v{Z}ilinskas, \textit{Simplicial Global Optimization} (Springer Briefs in Optimization. Springer, 2014).
%
%\bibitem{Ciegis}
%R.~\v{C}iegis, D.~Henty, B.~K\r{a}gstr\"om and J.~\v{Z}ilinskas, \textit{Parallel scientific computing and optimization: advances and applications}  (Springer, 2009). 
%
%\bibitem{Luque}
%G.~Luque and E.~Alba, \textit{Parallel genetic algorithms. Theory and real world applications} (Springer-Verlag, Berlin, 2011).
%
%\bibitem{Strongin2}
%R.~G.~Strongin, V.~P.~Gergel, V.~A.~Grishagin and K.~A.~Barkalov, \textit{Parallel computations for global optimization problems} (Moscow State University, Moscow, 2013) [In Russian].
%
%\bibitem{Sergeyev2}
%Ya.~D.~Sergeyev, R.~G.~Strongin, D.~Lera, \textit{Introduction to Global Optimization Exploiting Space-Filling Curves} (Springer Briefs in Optimization, Springer, 2013).
%
%\bibitem{Strongin3}
%R.~G.~Strongin, \textit{Numerical Methods in Multiextremal Problems (Information-Statistical Algorithms)} (Nauka, Moscow, 1978) [In Russian].
%
%\bibitem{Strongin4}
%R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global multidimensional optimization on parallel computer}, Parallel Computing \textbf{18} (11), 1259--1273 (1992).
%
%\bibitem{Sergeyev4}
%Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel asynchronous global search and the nested optimization scheme}, J. Comput. Anal. Appl., \textbf{3} (2), 123--145 (2001).
%
%\bibitem{Gergel1}
%V.~P.~Gergel and S.~V.~Sidorov, \textit{A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems}, Lecture Notes in Computer Science \textbf{9251}, 505--515 (2015).
%
%\bibitem{Gergel2}
%V.~Gergel, \textit{An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems}, in Proceedings of the Second International Conference on Mathematics and Computers in Sciences and in Industry (MCSI), Sliema, 13--18 (2015).
%
%\bibitem{Barkalov}
%K.~Barkalov, V.~Gergel and I.~Lebedev, \textit{Solving global optimization problems on GPU cluster}, AIP Conference Proceedings \textbf{1738}, 400006 (2016).
%
%\bibitem{Gergel3}
%V.~Gergel and E.~Kozinov, \textit{Efficient methods of multicriterial optimization based on the intensive use of search information}, Springer Proceedings in Mathematics and Statistics \textbf{197}, 27--45 (2017). 
%
%\bibitem{Gergel4}
%V.~Gergel and E.~Kozinov, \textit{Parallel computing for time-consuming multicriterial optimization problems}, Lecture Notes in Computer Science \textbf{10421}, 446--458 (2017).
%
%\bibitem{Gergel5}
%V.~Gergel, V.~Grishagin and A.~Gergel, \textit{Adaptive nested optimization scheme for multidimensional global search}, J. Glob. Optim.  \textbf{66} (1), 35--51 (2016).
%
%\bibitem{Lera}
%D.~Lera, Y.~D.~Sergeyev, \textit{Lipschitz and Holder global optimization using space-filling curves},  Appl. Numer. Math. \textbf{60} (1-2), 115--129 (2010).
%
%\bibitem{Grishagin1}
%V~A.~Grishagin. \textit{On convergence conditions for a class of global search algorithms}, in Proceedings of the 3-rd All-Union seminar ``Numerical methods of nonlinear programming'', Kharkov, 82--84 (1979) [In Russian].
%
%\bibitem{Grishagin2}
%V.~A.~Grishagin, Y.~D.~Sergeyev and R.~G.~Strongin, \textit{Parallel characteristic algorithms for solving problems of global optimization}, J. Glob. Optim. \textbf{10} (2), 185--206 (1997).
%
%\bibitem{Strongin5}
%R.~G.~Strongin, \textit{Algorithms for Multi-extremal Mathematical Programming Problems Employing the Set of Joint Space-filling Curves}, J. Glob. Optim. \textbf{2} (4), 357--378 (1992).
%
%\bibitem{Sysoyev}
%A.~Sysoyev, K.~Barkalov, V.~Sovrasov, I.~Lebedev and V.~Gergel, \textit{Globalizer -- A parallel software system for solving global optimization problems}, Lecture Notes in Computer Science \textbf{10421}, 492--499 (2017). 
%
%\bibitem{Sergeyev7}
%Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel Asynchronous Global Search and the Nested Optimization Scheme}, J. Comput. Anal. Appl. \textbf{3} (2), 123--145 (2001).
%
%\bibitem{Gaviano}
%M.~Gaviano, D.~Lera, D.~E.~Kvasov and Ya.~D.~Sergeyev, \textit{Software for generation of classes of test functions with known local and global minima for global optimization}, ACM Trans. Math. Softw. \textbf{29}, 469--480 (2003).
%
%\bibitem{Modorskii}
%V.~Y.~Modorskii, D.~F.~Gaynutdinova, V.~P.~Gergel and K.~A.~Barkalov, \textit{Optimization in design of scientific products for purposes of cavitation problems}, AIP Conference Proceedings \textbf{1738}, 400013 (2016).
%
%\bibitem{Gergel6}
%V.~P.~Gergel, M.~I.~Kuzmin, N.~A.~Solovyov and V.~A.~Grishagin, \textit{Recognition of surface defects of cold-rolling sheets based on method of localities}, International Review of Automatic Control \textbf{8} (1), 51--55 (2015).

\end{thebibliography}
\end{document}

### 2021-01-30 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
first version in Russian

@@
 
-\titlerunning{Short form of the title} % for running heads
+\titlerunning{Parallel global search algorithm with local tuning} % for running heads
 %\authorrunning{First-Author at al.} % for running heads
@@
 
-\title{Title of the Article,\\
-Broken into Lines}
+\title{Parallel Global Search Algorithm with Local Tuning\\ 
+for Solving Mixed-Integer Global Optimization Problems}
 % Splitting into lines is performed by the command \\
@@
 
-To prepare your manuscript, use REV\TeX{} package. The latest version can be downloaded at the project homepage \cite{RTeXHome}.
-The process of processing an article using REV\TeX{} is described in detail in the package manuals \cite{RTeX}. Great help in resolving technical
-issues on \TeX{} can be found in books \cite{texbook,L,GG,KD}.
-Use the ``House Style Guide: Version 2.0'' \cite{hsg} for more help on how to format article.
+В работе рассматриваются global optimization problems и методы их решения.
+Задачи глобальной оптимизации обладают значительной вычислительной трудоемкостью, т.к.
+глобальный оптимум является интегральной характеристикой решаемой задачи и требует исследования всей области поиска.
+Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (вообще говоря, неравномерного) в области параметров.
+Особый интерес представляют задачи, в которых часть параметров может принимать лишь целочисленные значения (mixed-integer global optimization problems), т.к.
+для них сложнее построить оценки оптимума по сравнению с непрерывными задачами.
 
-\section{First level Heading\protect\\
-broken into lines}
+Ситуация, когда некоторые из параметров характеризуются дискретностью (в частности, целочисленностью), является широко распространенной в прикладных оптимизационных задачах.
+При этом целочисленные параметры имеют, как правило, небольшое число значений и могут определять, например, марку используемого материала, вариант типовой компоновки деталей и т.п.
 
-There are three levels of Headings that are set by the commands
-\verb|\section|, \verb|\subsection|, and \verb|\subsubsection|
-(Chapter, subchapter, subsection). Split header into lines
-is done with the command \verb+\protect\\+.
-Second level heading as the title of article is written in accordance with the rules of capitalization \cite{cap}.
+Методам решения mixed-integer problems посвящена обширная литература (см., например, обзоры \cite{Burer,Boukouvala}).
+Известные детерминированные методы решения задач данного класса основаны, как правило, на подходах Branch-and-Bound \cite{Belotti} or Branch-and-Reduce \cite{Vigerske}.
+Также известен ряд метаэвристических и генетических алгоритмов, так или иначе основанных на идеях случайного поиска \cite{Deep,Schluter}.
 
-\subsection{Second Level Heading}
+%Ссылки на параллельные генетические алгоритмы - простота распараллеливания
+Вопросам распараллеливания генетических алгоритмов посвящены тысячи статей, т.к. данный класс алгоритмов допускает естественное распараллеливание.
 
-References in the text are by using the commands
-\verb+\cite{#1}+ or \verb+\onlinecite{#1}+. Label \verb+#1+
-can have a name consisting of both letters and numbers. In
-the bibliography section this link also has a label \verb+#1+ and starts the command \verb+\bibitem{#1}+.
+%Распараллеливание детерминированных алгоритмов - сложная задача
 
-With the command \verb+\cite{texbook}+ get a reference \cite{texbook},
-if you need to refer to several sources:
-\cite{texbook,L}, the curly brackets indicate references
-separated by commas: \verb+\cite{texbook,L}+. When
-use the command \verb|\onlinecite{#1}| link will not be
-in brackets: see~\onlinecite{texbook}, p.~8.
-If there is a link to the sources with serial numbers,
-e.g. [1,2,3,6,7,8], then print automatically following links will take
-[1--3,6--8].
+Отметим ключевые моменты, специфичные для построения детерминированных параллельных алгоритмов решения задач глобальной оптимизации.
+Организовать распараллеливание процесса поиска можно несколькими способами.
+Во-первых, можно распараллеливать процесс проведения одного испытания, т.е.
+вычисления функций задачи в одной точке области поиска.
+Данный путь будет давать ускорение, но является специфичным для каждой конкретной решаемой задачи.
+Во-вторых, можно распараллелить реализацию вычислительных правил алгоритма, обеспечивающих выбор точки очередного испытания.
+В этом случае способ распараллеливания будет зависеть от конкретного класса алгоритмов и, кроме того, часто эти правила достаточно просты и распараллеливать их нецелесообразно.
+В-третьих, можно организовать параллельное выполнение нескольких испытаний на одной итерации поиска.
+Этот подход и будет использоваться в данной работе, т.к.
+он характеризуется эффективностью (распараллеливается именно та часть вычислительного процесса, в которой выполняется основной объем вычислений) и общностью (применим для широкого класса алгоритмов глобальной оптимизации).
 
-\section{Standalone formulae}
+В данной работе нами предложен новый parallel метод решения mixed-integer problems, основанный на information-statistical approach к решению задач глобальной оптимизации \cite{Strongin2000,Strongin2013}.
 
-\subsection{Another second-level heading}
+В рамках данного подхода 
+\begin{itemize}
+	\item 
+	решение многомерных задач сводится  к решению эквивалентных им одномерных, cоответствующая редукция основана на использовании space-filling curves;
+	\item 
+	при решении constrained optimization problems каждое ограничение учитывается и обрабатывается отдельно, штрафные функции не используются;
+	\item
+	%локальное ускорение
+	\item 
+	распараллеливание процесса поиска осуществляется за счет одновременного вычисления на каждой итерации нескольких значений целевой функции в разных точках search domain.	
+\end{itemize}
 
-In \LaTeX, there are many ways to embed standalone formulas
-on the page and align them. By default, formulas always
-centered.
+Текст статьи, отражающей результаты проведенного исследования, построен следующим образом. 
+В Section 2 дано краткое описание описание схемы редукции размерности с использованием space-filling curves, а также описана индексная схема учета ограничений.
+Здесь же приведена формулировка параллельного индексного алгоритма для решения задач непрерывной глобальной оптимизации.
+В Section 3 изложен способ учета локальных свойств функций задачи, позволяющий ускорить сходимость алгоритма.
+В Section 4 предложен подход к обобщению параллельного индексного алгоритма с целью решения mixed-integer problems.
+Приведен способ, позволяющий свести решение mixed-integer problem к решению набора задач непрерывной оптимизации, которое может выполняться параллельно.
+Section 5 содержит результаты численных экспериментов.
+Здесь производится сравнение последовательной версии алгоритма с известными аналогами, а также показана эффективность параллельной версии алгоритма с локальной настройкой при решении серии multiextremal mixed-integer problem.
+Section 6 concludes the paper.
 
-\subsubsection{One-line formulas}
+\section{Parallel global optimization algorithm and dimensionality reduction}
 
-Below are examples of one-line equations:
-\begin{eqnarray}
-\chi_+(p)\alt{\bf [}2|{\bf p}|(|{\bf p}|+p_z){\bf ]}^{-1/2}
-\left(
-\begin{array}{c}
-|{\bf p}|+p_z\\
-px+ip_y
-\end{array}\right)\;,
-\\
-\left\{%
-\openone234567890abc123\alpha\beta\gamma\delta1234556\alpha\beta
-\frac{1\sum^{a}_{b}}{A^2}%
-\right\}%
-\label{one}.
+Рассмотрим задачу глобальной оптимизации вида
+\begin{eqnarray}\label{problem}
+&\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
+&D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.\label{D}
 \end{eqnarray}
-The second formula has the number~(\ref{one}), which is set by command
-\verb|\label{one}|. The first formula is assigned a number~(1), but
-it cannot be applied with the help of automatic links,
-as it has no label.
+Будем предполагать, что objective function $\varphi(y)$ (hereafter denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
+\[
+\left|g_i(y_1)-g_i(y_2)\right|\leq L_i\left\|y_1-y_2\right\|, \;1\leq i\leq m+1, \; y_1,y_2 \in D,\;
+\]
+with a priori unknown constants $L_i, \; 1 \leq i \leq m+1$.
 
-If the formula number is not necessary, use the environment
-\verb+\[+, \verb+\]+, (or \$...\$) which obtained the following formula:
-\[g^+g^+ \rightarrow g^+g^+g^+g^+ \dots ~,~~q^+q^+\rightarrow
-q^+g^+g^+ \dots ~. \] % You also can use the dollar sign ($).
+%Также будем предполагать, что функции задачи $g_i(y)$ определены и вычислимы только в тех точках обалсти $y \in D$, в которых выполняются ограничения
+%\begin{equation}\label{g_k}
+%g_k(y) \leq 0, \; 1 \leq k < i.
+%\end{equation}
 
-\subsubsection{Multiline formulae}
-
-Multiline formulas are typed using the environment
-eqnarray:
-\begin{eqnarray}
-{\cal M}=&&ig_Z^2(4E_1E_2)^{1/2}(l_i^2)^{-1}
-\delta_{\sigma_1,-\sigma_2}
-(g_{\sigma_2}^e)^2\chi_{-\sigma_2}(p_2)\nonumber\\
-&&\times
-[\epsilon_jl_i\epsilon_i]_{\sigma_1}\chi_{\sigma_1}(p_1),
-\end{eqnarray}
-\begin{eqnarray}
-\sum \vert M^{\rm viol}_g \vert ^2&=&g^{2n-4}_S(Q^2)~N^{n-2}
-(N^2-1)\nonumber \\
- & &\times \left( \sum_{i<j}\right)
-\sum_{\rm perm}
-\frac{1}{S_{12}}
-\frac{1}{S_{12}}
-\sum_\tau c^f_\tau~.
-\end{eqnarray}
-If the formula number is not necessary, then at the end of the row in front of the sign
-\verb|\\| you need to put the command \verb|\nonumber|. Never
-use one line command \verb|\nonumber| and
-\verb|\label{#1}|, as this may cause error in automatic
-the numbering of references.
-
-If you want to gain a few formulas without number,
-use the eqnarray environment* (the asterisk means the abolition
-numbering):
-\begin{eqnarray*}
-\sum \vert M^{\rm viol}_g \vert ^2&=&g^{2n-4}_S(Q^2)~N^{n-2}
-(N^2-1)\\
-& &\times \left( \sum_{i<j}\right)
-\left(
-\sum_{\rm perm}\frac{1}{S_{12}S_{23}S_{n1}}
-\right)
-\frac{1}{S_{12}}~.
-\end{eqnarray*}
-To add numbers to the formula manually, use the command
-\verb+\tag{#1}+, where \verb+#1+~--- the desired equation number.
-Here how is the formula with the number of~(\ref{eq:mynum}):
-\begin{equation}
-g^+g^+ \rightarrow g^+g^+g^+g^+ \dots ~,~~q^+q^+\rightarrow
-q^+g^+g^+ \dots ~. \tag{2.6$'$}\label{eq:mynum}
+Используя отображения типа Peano-Hilbert curve $y(x)$, отображающие отрезок $[0,1]$ на $N$-dimensional domain (\ref{D}),
+можно свести многомерную задачу условной минимизации в области $D$ к одномерной задаче условной минимизации на отрезке $[0,1]$
+\begin{equation}\label{problem1}
+\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
 \end{equation}
 
-When you enable single-line and multi-line formulas are surrounded by
-subequations, each formula `numbered" with a letter,
-as shown in equations~(\ref{mlett:1}) and (\ref{mlett:2}):
-\begin{subequations}
-\label{generallabel}
-\begin{equation}
-\left\{
-abc123456abcdef\alpha\beta\gamma\delta1234556\alpha\beta
-\frac{1\sum^{a}_{b}}{A^2}
-\right\},\label{mlett:1}
+Методы для построения чиcленных аппроксимаций Peano-Hilbert curve $y(x)$ (называемых \textit{evolvent}) рассмотрены в \cite{Strongin2013}.
+These evolvents are fractals generated by an iterative process, that fill in the hypercube $D$ with accuracy $2^{-M}$, where integer $M>0$ is the evolvent construction parameter.
+Примеры разверток с разными значениями $M$  для размерности $N=3$ приведены на Fig.~\ref{evolvents}.
+
+\begin{figure}
+\begin{minipage}{0.49\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (b)}
+\end{minipage}
+\caption{Evolvents in three dimensions with (a) $M=3$ and (b) $M=4$}
+\label{evolvents}
+\end{figure}
+
+Задачу (\ref{problem1}) мы будем рассматривать в предположении, что функции $g_i(y(x))$ определены и вычислимы лишь на соответствующих множествах 
+\[
+Q_1=[0,1], \; Q_{i+1}=\left\{x \in Q_i : g_i(y(x)) \leq 0 \right\}, \; 1 \leq i \leq m.
+\]
+
+Эти условия позволяют ввести классификацию точек $x \in [0,1]$ в соответствии с числом $\nu (x)$ ограничений, вычисленных в этой точке.
+The \textit{index} $\nu(x)$ также можно определить с помощью условий
+\begin{equation}\label{nu}
+g_i(y(x)) \leq 0, \; 1 \leq i < \nu, \; g_\nu(y(x))>0,
 \end{equation}
-\begin{eqnarray}
-{\cal M}=&&ig_Z^2(4E_1E_2)^{1/2}(l_i^2)^{-1}
-(g_{\sigma_2}^e)^2\chi_{-\sigma_2}(p_2)\nonumber\\
-&&\times
-[\epsilon_i]_{\sigma_1}\chi_{\sigma_1}(p_1).\label{mlett:2}
+где последнее неравенство исключается из рассмотрения, если $\nu=m+1$.
+
+Указанная схема редукции размерности сопоставляет многомерной задаче с липшицевой минимизируемой функцией и липшицевыми ограничениями одномерную задачу, в которой соответствующие функции удовлетворяют uniform H{\"o}lder condition (see \cite{Strongin2013}), i.e.,
+\[
+\left|g_i(y(x_1))-g_i (y(x_2))\right| \leq H_i \left|x_1-x_2 \right|^{1/N}, \; x_1,x_2\in [0,1], \; 
+1\leq i \leq m+1,
+\]
+где $N$ есть размерность исходной многомерной задачи, а H{\"o}lder coefficients $H_i$ связаны с Lipschitz constants $L_i$ исходной задачи соотношениями $H_i \leq 2L_i \sqrt{N+3}$.
+
+Таким образом, \textit{a trial} в точке $x^k \in [0,1]$, выполненное на $k$-й итерации алгоритма, будет состоять из следующих действий:
+\begin{itemize}
+	\item определить \textit{image} $y^k=y(x^k)$ в соотвествии с отображением $y(x)$;
+	\item вычислить значения $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from (\ref{nu}). 
+\end{itemize}
+Пару значений 
+\begin{equation} \label{trial_result}
+ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} 
+\end{equation}
+порождаемую испытанием в точке $x^k \in [0,1]$, будем называть \textit{trial outcome}.
+
+%эффективный алгоритм
+%исходная информация - параллельность
+Пусть в нашем распоряжении имеется $p$ вычислительных элементов (ядер или процессоров), которые будут использоваться для параллельного проведения $p$ испытаний одновременно.
+На первой итерации метода $p$ испытаний проводятся параллельно в произвольных различных точках $x^i\in(0,1)$, $1\leq i \leq p$.
+
+Пусть выполнено $n\geq 1$  итераций метода, в процессе которых были проведены испытания в $k=np$ точках $x^i, 1\leq i \leq k$.
+Тогда точки $x^{k+1},...,x^{k+p}$  поисковых испытаний $(n+1)$-ой итерации определяются в соответствии со следующими правилами.
+
+%Проверить перевод!
+\begin{enumerate}
+\item 
+Перенумеровать точки $x^1,...,x^k$ предшествующих испытаний нижними индексами в порядке увеличения значений координаты
+\begin{equation}\label{Eq:17}
+0=x_0<x_1<...<x_i<...<x_k<x_{k+1}=1,
+\end{equation}
+и сопоставить им значения $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, из (\ref{trial_result}), вычисленные в этих точках; точки $x_0=0$ and $x_{k+1}=1$ введены дополнительно; значения $z_0$ и $z_{k+1}$ не заданы.
+\item
+Провести классификацию номеров  $i,1\leq i \leq k$, точек испытаний из  (\ref{Eq:17}) по числу ограничений задачи, выполняющихся в этих точках, путем построения множеств
+\begin{equation}\label{Eq:18}
+I_\nu = \left\{i: 1 \leq i \leq k,\ \nu = \nu(x_i)\right\},\ 1 \leq \nu \leq m+1,
+\end{equation}
+содержащих номера всех точек  $x_i,1\leq i \leq k$, имеющих индексы, равные одному и тому же значению $\nu$.
+Граничные точки $x_0=0$ and $x_{k+1}=1$ интерпретируются как имеющие нулевые индексы, и им сопоставляется дополнительное множество $I_0={0,k+1}$.
+
+Определить максимальное значение индекса
+\begin{equation}\label{Eq:19}
+V=\max \left\{\nu = \nu(x_i), \ 1\leq i \leq k\right\}.
+\end{equation}
+\item
+For all values of $\nu, \ 1\leq \nu \leq m+1$, calculate the values  
+\begin{equation}\label{Eq:20}
+\mu_\nu = \max \left\{ \frac{\left|z_i-z_j\right|}{\left(x_i-x_j\right)^{1/N}} : i,j \in I_\nu, j<i\right\}.
+\end{equation}
+Если множество  $I_\nu$ содержит менее двух элементов или если $\mu_\nu$ from (\ref{Eq:20}) оказывается равным нулю, то использовать значение $\mu_\nu=1$.
+\item
+Для всех непустых множеств  $I_\nu$, $1 \leq \nu \leq m+1$, вычислить значения
+\begin{equation}\label{Eq:21}
+  z^\ast_\nu =  
+   \begin{cases}
+    -\epsilon_\nu,  \nu < V, \\
+    \min{\left\{g_\nu(x_i):i\in I_\nu\right\}}, \nu = V,
+   \end{cases}
+\end{equation}
+где $V$ является максимальным значением индекса, а вектор
+\begin{equation}\label{Eq:22}
+\epsilon _R=\left(\epsilon_1,...,\epsilon_m\right),
+\end{equation}
+с положительными компонентами (называемый также \textit{reserve vector}) является параметром алгоритма.
+\item
+Для каждого интервала  $(x_{i-1},x_i)$,$1 \leq i \leq k+1$, вычислить \textit{характеристику} $R(i)$: 
+\begin{eqnarray}\label{R}
+&R(i)=\Delta_i+ \frac{(z_i-z_{i-1})^2}{(r_\nu\mu_\nu)^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu=\nu(x_{i-1})=\nu(x_i),\nonumber \\
+&R(i)= 2\Delta_i-4\frac{z_i-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu(x_{i-1})<\nu(x_i)=\nu,\\
+&R(i)= 2\Delta_i-4\frac{z_{i-1}-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu = \nu(x_{i-1})>\nu(x_i).\nonumber
 \end{eqnarray}
-\end{subequations}
-If you put the label right after the \verb+\begin{subequations}+,
-it can be used further as a reference to all the equations in this
-environment. For example, you can refer to
-equation~(\ref{generallabel}) this example.
+где $\Delta_i=(x_i-x_{i-1})^{1/N}$, а значения $r_\nu>1, 1\leq\nu\leq m+1$, являются параметрами алгоритма.
+\item
+Упорядочить характеристики $R(i)$, $1\leq i \leq k+1$, по убыванию  	
+\begin{equation}\label{Eq:23}
+R(t_1)\geq R(t_2)\geq ... \geq R(t_{k})\geq R(t_{k+1})
+\end{equation}
+и выбрать $p$ интервалов с номерами $t_j, 1\leq j \leq p$, которым соответствуют наибольшие характеристики.
+\item
+Провести параллельно $p$ новых испытаний в точках $x^{k+j}, 1 \leq j \leq p$, вычисленных по формулам
+\begin{eqnarray*}
+& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}, \; \nu(x_{t_j-1})\neq \nu(x_{t_j}), \\
+& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}- \frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \; \nu(x_{t_j-1})=\nu(x_{t_j})=\nu. \\
+\end{eqnarray*}
 
-To set multi-line formulas you can use the environment
-multline, gather and align. The multline environment is good to use
-for a set of standalone long formulas that do not fit on
-one line:
-\begin{multline}
-\int_{a_1}^{a_2} f(x)\,dx+\int_{a_2}^{a_3} f(x)\,dx
-+\dots+\int_{a_{n-1}}^{a_n} f(x)\,dx\\
-+\int_{a_1}^{a_2} g(x)\,dx+\int_{a_2}^{a_3} g(x)\,dx
-+\dots+\int_{a_{n-1}}^{a_n} g(x)\,dx\\
-+\int_{a_1}^{a_2} h(x)\,dx+\int_{a_2}^{a_3} h(x)\,dx
-+\dots+\int_{a_{n-1}}^{a_n} h(x)\,dx\\
-=\int_{a_1}^{a_n} f(x)+g(x)+h(x)\,dx.
-\end{multline}
-This formula is automatically numbered, if the formula number is not
-need, you have to use the environment multline*.
+\end{enumerate}
 
-Environment gather centers included in the formula:
-\begin{gather}
-\int_{a_1}^{a_2} f(x)\,dx+\int_{a_2}^{a_3} f(x)\,dx
-+\dots+\int_{a_{n-1}}^{a_n} f(x)\,dx\\
-+\int_{a_1}^{a_2} g(x)\,dx+\int_{a_2}^{a_3} g(x)\,dx
-+\dots+\int_{a_{n-1}}^{a_n} g(x)\,dx\notag\\
-+\int_{a_1}^{a_2} h(x)\,dx+\int_{a_2}^{a_3} h(x)\,dx
-+\dots+\int_{a_{n-1}}^{a_n} h(x)\,dx\\
-=\int_{a_1}^{a_n} f(x)+g(x)+h(x)\,dx.
-\end{gather}
-Each line is automatically numbered, if the line number
-it is not necessary, before \verb+\\+ in this line, you need to put the command
-\verb+\notag+. When you use the environment gather* formula
-will not be numbered.
+Алгоритм прекращает работу, если по крайней мере для одного из номеров $t_j, 1\leq j \leq p$ выполняется условие $\Delta_{t_j}\leq \epsilon$; здесь $\epsilon>0$ есть заданная точность решения задачи.
 
-The align environment allows you to align formulas on your
-discretion:
-\begin{align}
-\int_{a_1}^{a_2} f(x)\,dx &+\int_{a_2}^{a_3} f(x)\,dx
-+\dots+\int_{a_{n-1}}^{a_n} f(x)\,dx\notag\\
-&+\int_{a_1}^{a_2} g(x)\,dx +\int_{a_2}^{a_3} g(x)\,dx
-+\dots+\int_{a_{n-1}}^{a_n} g(x)\,dx\\
-&+\int_{a_1}^{a_2} h(x)\,dx+\int_{a_2}^{a_3} h(x)\,dx
-+\dots+\int_{a_{n-1}}^{a_n} h(x)\,dx\notag\\
-&=\int_{a_1}^{a_n} f(x)+g(x)+h(x)\,dx.
-\end{align}
-Read more about working with these environments can be found in the book
-\cite{GG}.
+A detailed description of the algorithm convergence theory is presented in \cite{Strongin2000,Strongin2013}.
 
-\section{Variables}
-For the writing of definitions, theorems, lemmas and their proofs use the following variables. If necessary, you can add your variables in the sample.
+%комментарий по поводу r и eps-резервов
 
-\begin{definition}\label{D:1}
-Define ...
-\end{definition}
+%комментарий про различные схемы распараллеливания - со ссылками на наши работы
 
-\begin{lemma}\label{L:1}
-If ...
-\end{lemma}
+\section{Local tuning for global optimization algorithm}
 
-\begin{theorem}\label{Th:1}
-Let ...
-\end{theorem}
-\begin{proof}
-Consider...
-\end{proof}
+Одно из важных свойств рассматривамых задач заключается в том, что во многих случаях поведение функций в задаче многоэкстремальной оптимизации является неоднородным в разных подобластях поиска.
+В некоторых подобластях значения функций могут изменяться достаточно быстро (что будет соответствовать большим значениям константы Липшица в таких подобластях), в других подобластях значения могут изменяться более плавно.
+Как результат, построение единых оценок характеристик оптимизируемых функций для всей области поиска может не полностью соответствовать ее поведению в конкретной подобласти.
 
-\section{Figures and tables}
-REV\TeX~4 will automatically number sections, equations, tables, and
-drawings. Figure captions are made after the image and the signature
-tables~--- before a table, as shown in the examples at the end. In
-tables footnotes \footnote{Footnotes in tables, you can try
-to do this manually.} not working.
+Среди возможных подходов, учитывающих неоднородность поведения оптимизируемых функций в разных подобластях области поиска, можно выделить направление, связанное с адаптивным прогнозированием положения экстремума на основе накопленной поисковой информации в предположении слабой многоэкстремальности задачи.
 
-Graphics and diagrams should be monochrome in black-and-white color and saved in EPS format.
-%You can include a picture in the article using the figure environment:
-%\begin{figure}[h]
-%\setcaptionmargin{5mm}
-%%\onelinecaptionsfalse % if the caption is multiline
-%\onelinecaptionstrue  % if the caption is one-line
-%\includegraphics[width=0.85\textwidth]{deform.eps}
-%\captionstyle{normal}\caption{Please write your figure caption here.}\label{fig:1}
-%\end{figure}
+%Априорная информация, имеющаяся при решении прикладных многоэкстремальных задач, помимо факта ограниченности разделенных разностей функций задачи может содержать дополнительные характеристики экстремума, использование которых ускоряет процесс поиска.
+К числу таких характеристик можно отнести некоторый априорный прогноз расположения оптимальной точки (основанный, например, на наличии прототипа оптимизируемого объекта).
+%Если же такой априорный прогноз отсутствует,
 
-\begin{table}[!h]
-\setcaptionmargin{0mm}
-\onelinecaptionsfalse
-\captionstyle{flushleft}
-\caption{ For the insertion of tables, the table environment is used,
- the signatures to the tables are made in the same way as the captions
- for the figures. This is an example of a table whose multi-line name
- is decorated with the \textbf{caption2} package.}
-\bigskip
-\begin{tabular}{|c|c|c|c|c|c|c|c|}
+А именно, можно построить адаптивные правила прогнозирования, основанные на предположении о том, что местонахождение глобального оптимума более вероятно в окрестностях тех точек из ряда $x_i, 1 \leq i \leq k,$ которым соответствуют малые значения нормированных разностей $(z_i-z_\nu^*)/\mu_\nu$
+где $z_i$ из (\ref{trial_result}), $\mu_\nu$ из (\ref{Eq:20}), а $z_\nu^*$ из (\ref{Eq:21}) .
+
+Принятая нами гипотеза правдоподобна либо в случае, когда число осуществленных итераций достаточно велико (т.к.
+алгоритм при выполнении условий сходимости порождает последовательность испытаний, плотную лишь в точках глобального минимума), либо в случае, когда задача является унимодальной (т.е.
+когда поиск в окрестности лучшего текущего приближения позволяет достичь улучшения, и такие последовательные улучшения ведут к решению).
+Первый случай соответствует заключительной фазе поиска, и тогда использование адаптивного прогноза можно интерпретировать как способ ускорения уточнения решения.
+Второй случай фактически включает в себя предположения о слабой многоэкстремальности задачи.
+
+Конкретным примером новых формул для вычисления характеристик $R'(i)$, включающих в себя адаптивый прогноз расположения оптимума, являются выражения
+\begin{eqnarray}\label{lR}
+& R'(i) = \frac{R(i)}{\sqrt{(z_i-z_\nu^*)(z_{i-1}-z_\nu^*)}/\mu_\nu+1.5^{-q}}, \; \nu(x_{i-1})=\nu(x_{i}),\nonumber\\
+& R'(i) = \frac{R(i)}{(z_i-z_\nu^*)/\mu_\nu+1.5^{-q}}, \; \nu(x_{i-1})<\nu(x_{i}),\\
+& R'(i) = \frac{R(i)}{(z_{i-1}-z_\nu^*)/\mu_\nu+1.5^{-q}}, \; \nu(x_{i-1})>\nu(x_{i}),\nonumber
+\end{eqnarray}
+где $R(i)$ из (\ref{R}) , а целочисленный параметр $q$ характеризует «степень локальности» алгоритма. При $q=1$ поиск имеет «глобальный», а при $q>>1$ -- «локальный» характер. 
+%Формулы (\ref{localR}) следует использовать в правилах алгоритма на шаге вместо
+
+Управление параметром $q$, входящим в (\ref{lR}), позволяет учитывать различные предположения о характере экстремума.
+Например, одной из стратегий является чередование с заданной частотой больших и малых значений $q$, что соответствует смеси локального и глобального поиска, которую можно интерпретировать как сочетание процедуры локального уточнения текущего решения и процедуры глобального обновления этого решения.
+При этом в течение первых итераций рекомендуется использовать значение $q=1$, т.к.
+накопленная на них поисковая информация обеспечит адекватность формул (\ref{lR}) при последующем локальном уточнении.
+
+%Отметим, что на скорость сходимости алгоритмов глобальной оптимизации существенно влияет константа Липшица,
+%поэтому столь важным является вопрос ее корректной оценки.
+Заниженная оценка истинной константы Липшица L может привести к потере глобального решения.
+В то же время завышенная оценка константы L, не соответствующая истинному поведению функции, влечет за собой медленную сходимость алгоритма к точке глобального минимума.
+%Использование только глобальной информации о поведении целевой функции в ходе ее минимизации может значительно замедлить сходимость алгоритма к точке глобального минимума, т.к.
+могут существовать зоны области поиска, которые находятся далеко от точки глобального минимума, но локальные константы Липшица в которых существенно меньше глобальной.
+%
+%В работах [47, 140, 156] был предложен метод локальной настройки, преодолевающий указанные сложности и позволяющий адап-
+%тивно оценивать локальные константы Липшица на всей области поиска. Балансирование локальной и глобальной информации в про-
+%цессе поиска глобального минимума позволяет существенно ускорить работу алгоритма глобальной оптимизации.
+%
+%
+%Суть предложенного в подхода состоит в следующем.
+В алгоритме глобального поиска предлагается использовать не единую для всей области поиска оценку константы Липшица, а строить ее оценки для каждого поискового интервала в отдельности.
+Поскольку получение таких оценок может происходить с достаточно большими погрешностями, то при нахождении локальных (интервальных) оценок должна учитываться и общая (глобальная) оценка константы Липшица.
+%
+%Приведем алгоритмическое описание предлагаемого подхода.
+%Предположим, что параллельный алгоритм глобального поиска выполнил $k\geq$ итераций поиска и получены точки испытаний из (ссылка).
+%
+%Оценка константы Липшица   интервала (xi1,xi), 1<i<k, вычисляется в соот-ветствии со следующими выражениями:
+
+\section{Parallel algorithm for mixed-integer problems}
+
+Рассмотрим теперь случай, когда аргумент функций задачи содержит две компоненты: вектор $y$, принадлежащий гиперинтервалу $D$, и вектор $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е.
+\begin{eqnarray}\label{problem_i}
+& \min{\left\{ g_{m+1}(y,u):y\in D, \; g_i(y,u)\leq 0, \; 1 \leq i \leq m\right\}},\\
+& D=\left\{a_j\leq y_j \leq b_j, \; 1\leq j \leq N \right\}.\nonumber
+\end{eqnarray}
+
+Такие конечные наборы могут характеризовать, например, вариант материала, из которого создается объект, геометрические размеры или другие величины, которые могут принадлежать стандартному дискретному множеству, и т.п.
+
+Занумеруем целыми числами $s, 1\leq s \leq S,$ все возможные значения вектора $u$, т.е. сопоставим каждому рассматриваемому значению $s$ вектор $u_s$. 
+Тогда рассматриваемая задача может быть записана в виде 
+\begin{eqnarray}\label{problem_is}
+& \min_{s\in\{1,...,S\}}\left(\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_i(y,u_s)\leq 0, \; 1 \leq i \leq m\right\}}\right),\\
+& D=\left\{ a_j\leq y_j \leq b_j, \; 1 \leq j\leq N \right\}.\nonumber 
+\end{eqnarray}
+
+Используя схему редукции размерности с помощью Hilbert curve $y(x), x\in [0,1]$ можно сопоставить каждой задаче минимизации по $y$ одномерную задачу минимизации
+\[
+ \min{\left\{ g_{m+1}(y(x),u_s):x \in [0,1], \; g_i(y(x),u_s)\leq 0, \; 1 \leq i \leq m\right\}}, s\in\{1,...,S\}.
+\]
+
+Рассмотрим теперь соответствие 
+\[
+Y(x)=y(x-E(x)), \; x\in[0,S],
+\]
+отображающее любую точку интервала [0,S] на область $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функции 
+\[
+f_i(x) = g_i(Y(x),u_{E(x)+1}), x\in[0,S],
+\]
+имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_k = i, 1\leq i \leq S-1$.
+Значение  $z_k = g_\nu(y(x_k))$ из (\ref{trial_result}) в этих точках будем считать неопределенным, а значения индекса -- равным 0, т.е. $\nu(x_k) = 0$ .
+
+Используя введенные обозначения можно переформулировать исходную задачу как
+\begin{equation}\label{problem_is1}
+\min \left\{f_{m+1}(x): x \in [0,S], \; f_i(x) \leq 0, \; 1 \leq i \leq m\right\}.
+\end{equation}
+
+Для иллюстрации на Fig. \ref{fig:1}  изображены графики функций, соответствующих задаче с одним непрерывным и одним бинарным параметром.
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.7\textwidth]{fig1.jpg}
+    \caption{Mixed-integer global optimization problem}
+    \label{fig:1}
+\end{figure}
+
+Применяя к решению задачи (\ref{problem_is1}) параллельный индексный алгоритм с локальной настройкой, мы найдем решение задачи (\ref{problem_i}).
+При этом основная часть испытаний будет проведена в той подзадаче, решение которой соответствует решению исходной задачи (\ref{problem_i}).
+В остальных подзадачах будет проведена лишь незначительная часть испытаний, т.к.
+решения данных подзадач являются локально-оптимальными по отношению к решению $s$-й подзадачи.
+Сказанное подтверждается Fig.
+\ref{fig:1}, где штрихами обозначены точки испытаний, выполненных при решении данной задачи.
+
+Таким образом, мы сформировали Mixed-Integer Global Search Algorithm (MIGSA), основанный на сведении задачи mixed-integer non-convex optimization problem к одновременно решаемому набору непрерывных задач.
+Можно доказать, что условия сходимости данного алгоритма будут следовать из условий сходимости его прототипа (GSA).
+
+Предложенная схема решения mixed-integer global optimization problems основана на параллельном индексном методе для решения задач непрерывной оптимизации и не является ориентированной на конкретное вычислительное устройство.
+Одной из ключевых операций здесь является параллельное проведение нескольких испытаний в различных точках области поиска (см.
+шаг 7 алгоритма), которое может быть реализовано как на CPU (с использованием OpenMP и/или MPI), так и на GPU (с использованием CUDA).
+
+%здесь - сослаться на распараллеливание с использованием блочной многошаговой схемы.
+
+\section{Results of experiments}
+
+Первая серия экспериментов была проведена с использованием последовательной версии предложенного алгоритма с целью его сравнения с известными методами аналогичного назначения.
+
+Предложенный в данной работе алгоритм MIGSA сравнивался с genetic algorithm for solving mixed-integer optimization problems, реализованным в Matlab Global Optimization Toolbox.
+В Table \ref{tab:1} приведено число испытаний, потребовавшееся для решения известных тестовых mixed-integer problems данными методами.
+Для обоих методов была использована одинаковая точность поиска решения $10^{-2}$.
+Все вычислительные эксперименты были проведены на компьютере с процессором Intel Core i5-7300 2.5 GHz и 8 Gb RAM.
+Результаты экспериментов показывают превосходство метода MIGSA как по числу итераций, так и по времени работы.
+
+\begin{table}
+	\caption{Сравнение эффективности методов MIGSA и GA}
+	\label{tab:1}
+	\center
+	\begin{tabular}{|c|c|c|c|c|}
 		\hline
- &$r_c$ (\AA)&$r_0$ (\AA)&$\kappa r_0$&
- &$r_c$ (\AA) &$r_0$ (\AA)&$\kappa r_0$\\
+		%\noalign{\smallskip}		
+	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c|}{ GA } &  \multicolumn{2}{c|}{MIGSA} \\
+		%\noalign{\smallskip} 
+		\cline{2-3} \cline{4-5} 
+		%\noalign{\smallskip}
+		 & $k$ & $t$ &  $k$ & $t$  \\
+		%\noalign{\smallskip} 
 		\hline 
-Cu& 0.800 & 14.10 & 2.550 &Sn
-& 0.680 & 1.870 & 3.700 \\
-Ag& 0.990 & 15.90 & 2.710 &Pb
-& 0.450 & 1.930 & 3.760 \\
-Au& 1.150 & 15.90 & 2.710 &Ca
-& 0.750 & 2.170 & 3.560 \\
-Mg& 0.490 & 17.60 & 3.200 &Sr
-& 0.900 & 2.370 & 3.720 \\
-Zn& 0.300 & 15.20 & 2.970 &Li
-& 0.380 & 1.730 & 2.830 \\
-Cd& 0.530 & 17.10 & 3.160 &Na
-& 0.760 & 2.110 & 3.120 \\
-Hg& 0.550 & 17.80 & 3.220 &K &  1.120 & 2.620 & 3.480 \\
-Al& 0.230 & 15.80 & 3.240 &Rb & 1.330 & 2.800 & 3.590 \\
-Ga& 0.310 & 16.70 & 3.330 &Cs & 1.420 & 3.030 & 3.740 \\
-In& 0.460 & 18.40 & 3.500 &Ba & 0.960 & 2.460 & 3.780 \\
-Tl& 0.480 & 18.90 & 3.550 & & & & \\[1mm]
+		%\noalign{\smallskip}
+		 Problem 2 \cite{Floudas}&	481 &	0.0601 & 	417 &	0.04 \\
+		 Problem 6 \cite{Floudas}&	641 &	0.0510 & 	118 &	0.001 \\
+		 Problem 1 \cite{Deep}   &	481 &	0.1378 & 	66 &	0.0007 \\
+		 Problem 2 \cite{Deep}   &	481 &	0.0473 & 	57 &	0.0006 \\
+		 Problem 7 \cite{Deep}   &	841 &	0.0736 &  372	 &	0.017 \\
+		%\noalign{\smallskip}
 		\hline
@@
 
-\begin{table}[!htb]
-\setcaptionmargin{0mm}
-\onelinecaptionstrue
-\captionstyle{flushleft}
-\caption{The name of this table is --- one-line.}
-\bigskip
-\begin{tabular}{|c|c|c|c|c|c|c|}
-  \hline
-    & 1 & 2 & 3 & 4 & 5 & 6\\
-  \hline
-  1 & 1 & 2 & 3 & 4 & 5 & 6\\
-  2 & 2 & 4 & 6 & 8 & 10 & 12\\
-  3 & 3 & 6 & 9 & 12 & 15 & 18\\[1mm]
-  \hline
-\end{tabular}
-\end{table}
+%Последующие серии экспериментов были выполнены для оценки ускорения параллельной версии предложенного алгоритма MIPIA с использованием как CPU, так и GPU.
+В данных экспериментах решалась серия из 100 тестовых mixed-integer problems, сгенерированных случайных образом.
 
-\section{Bibliography}
-The following rules apply for references to books:
+Известным генератором тестовых задач непрерывной многоэкстремальной оптимизации является GKLS-генератор \cite{Gaviano}.
+Он позволяет генерировать тестовые функции заданной размерности с заданными свойствами: известным числом локальных минимумов, известной областью притяжения глобального решения, и т.д.
+%This generator of multiextremal functions is often used for the investigations of the global optimization algorithms \cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
 
-\begin{itemize}
-  \item Authors initials go before the surname with a space between the initials. Use and between the last two authors. All authors listed in the original reference should be cited.
-  \item The title of the book is written in italics.
-If the book is originally published in Russian, only the English translation of the title is cited and after citation the original language may be indicated in square brackets, e.g.: [in Russian].
-  \item Examples of bibliographic references can be found at the end of this page % (i.e. article \cite{ex}, PhD thesis \cite{maguire-76}).
-\end{itemize}
+In Fig.~\ref{example} (a) and (b), the contour plots of two-dimensional GKLS functions are presented.
+Figures also shows the points of the trials performed by the method until the required accuracy $\epsilon=10^{-2}$ was achieved.
+
+%\begin{figure}
+%\begin{minipage}{0.5\linewidth}
+%\center{\includegraphics[width=1.0\linewidth]{GKLS-4.png} \\ (a)}
+%\end{minipage}
+%\hfill
+%\begin{minipage}{0.5\linewidth}
+%\center{\includegraphics[width=1.0\linewidth]{GKLS-6.png} \\ (b)}
+%\end{minipage}
+%\caption{Solving a two-dimensional problem using the index algorithm}
+%\label{example}
+%\end{figure}
+
+В проведенных экспериментах GKLS-генератор использовался как основа для построения mixed-integer problems. Правила, позволяющие генерировать тестовые задачи подобного вида, заключаются в следующем.
+
+\begin{enumerate}
+	\item С помощью генератора GKLS порождается непрерывная многоэкстремальная функция $\psi(y), \; y\in D, \; D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\}$.
+Глобальный минимум данной функции достигается в известной точке $y'=(y'_1,...,y'_N)$ и равен $\psi'=\psi(y')=-1$.
+	\item Генерируется convex mixed-integer function 
+	\[
+			\mu(y,u) = 2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} \right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right],
+	\]
+	где 
+	\begin{eqnarray*}
+	& y\in D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\} \subset R^N, \\
+	& u\in U = \left\{ u_j \in  \left\{a_j, ..., b_j \right\}, 1\leq j \leq M \right\};
+	\end{eqnarray*}
+	т.е. данная функция имеет $N$ непрерывных и $M$ дискретных параметров и достигает минимального значения в точке $(y',b)$.
+	\item Вычисляется коэффициент 
+	\[
+	C = 4 - \min_{y,u} \left\{ \mu(y,u) \right\}, \; y\in D, u \in U.
+	\]
+	Минимум вогнутой функции $-\mu(y,u)$ достигается в одной из угловых точек области поиска, поэтому при размерности задачи порядка 10 его вычисление может быть выполнено методом перебора.
+	\item Формируется многоэкстремальная mixed-integer function 
+	\[
+	\varphi(y,u) = \left(\psi(y) + \sum_{j=1}^M{u_j}\right)\left(C - \mu(y,u)\right).
+	\]
+	По построению $\varphi(y,u)$  будет принимать минимальное значение в точке $(y',b)$.
+
+\end{enumerate}
+
+In the problems generated in our experiments, there were 5 discrete and 6 continuous parameters 
+	\begin{eqnarray*}
+	& y\in D = \left\{ -1 \leq y_j\leq 1, 1\leq j \leq 6 \right\} \subset R^6, \\
+	& u\in U = \left\{ u_j \in  \left\{-1, -1/3, 0, 1/3, 1 \right\}, 1\leq j \leq 5 \right\}.
+	\end{eqnarray*}
 
@@
 
+\bibitem{Burer}
+S.~Burer and A.~N.~Letchford, \textit{Non-convex mixed-integer nonlinear programming: A survey}, Surveys in Operations Research and Management Science \textbf{17}, 97--106 (2012).
+
+\bibitem{Boukouvala}
+F.~Boukouvala, R.~Misener and C.~A.~Floudas, \textit{Global optimization advances in Mixed-Integer Nonlinear Programming, MINLP, and Constrained Derivative-Free Optimization, CDFO}, European J.
+Oper.
+Res.
+\textbf{252}, 701--727 (2016).
+
+\bibitem{Belotti}
+P.~Belotti, J.~Lee, L.~Liberti, F.~Margot and A.~W\"achter, \textit{Branching and bounds tightening techniques for non-convex MINLP}, Optim. Method. Softw. \textbf{24}(4-5), 597--634 (2009).
+
+\bibitem{Vigerske}
+S.~Vigerske and A.~Gleixner, \textit{SCIP: global optimization of mixed-integer nonlinear programs in a branch-and-cut framework}, Optim. Method. Softw. \textbf{33}(3), 563--593 (2018).
+
+\bibitem{Deep}
+K.~Deep, K.~P.~Singh, M.~L.~Kansal and C.~Mohan, \textit{A real coded genetic algorithm for solving integer and mixed integer optimization problems}, Appl.
+Math.
+Comput.
+\textbf{212}(2), 505--518 (2009).
+
+\bibitem{Schluter}
+M.~Schl\"uter, J.~A.~Egea and J.~R.~Banga, \textit{Extended ant colony optimization for non-convex mixed integer nonlinear programming}, Computers and Operations Research \textbf{36}(7), 2217--2229 (2009).
+
+\bibitem{Strongin2000}
+R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global optimization with non-convex constraints.
+Sequential and parallel algorithms} (Kluwer Academic Publishers, Dordrecht, 2000, 2nd ed.
+2013, 3rd ed.
+2014).
+
+\bibitem{Strongin2013}
+R.~G.~Strongin, V.~P.~Gergel, V.~A.~Grishagin and K.~A.~Barkalov, \textit{Parallel computations for global optimization problems} (Moscow State University, Moscow, 2013) [In Russian].
+
+\bibitem{Floudas}
+C.~A.~Floudas and M.~P.~Pardalos,  \textit{Handbook of Test Problems in Local and Global Optimization}. (Springer, 1999).  %; DOI: 10.1007/978-1-4757-3040-1
+
 %\bibitem{Floudas}
@@
 %
-%\bibitem{Strongin1}
-%R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global optimization with non-convex constraints.
-Sequential and parallel algorithms} (Kluwer Academic Publishers, Dordrecht, 2000, 2nd ed.
-2013, 3rd ed.
-2014).
 %
@@
 %G.~Luque and E.~Alba, \textit{Parallel genetic algorithms. Theory and real world applications} (Springer-Verlag, Berlin, 2011).
-%
-%\bibitem{Strongin2}
-%R.~G.~Strongin, V.~P.~Gergel, V.~A.~Grishagin and K.~A.~Barkalov, \textit{Parallel computations for global optimization problems} (Moscow State University, Moscow, 2013) [In Russian].
 %

### 2021-02-01 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
Update text in Russian

@@
 
-В работе рассматриваются global optimization problems и методы их решения.
+В работе рассматриваются global optimization problems и параллельные методы их решения.
 Задачи глобальной оптимизации обладают значительной вычислительной трудоемкостью, т.к.
@@
 При этом целочисленные параметры имеют, как правило, небольшое число значений и могут определять, например, марку используемого материала, вариант типовой компоновки деталей и т.п.
+При этом целевая функция и ограничения обычно заданы не явно, а в виде некоторого алгоритма вычисления их значений в точках области поиска, т.е.
+представлять собой “black-box”-функции.
+Процесс проверки ограничений и вычисления значения целевой функции в некоторой точке (называемый в дальнейшем trial в этой точке) требует проведения численного моделирования и является трудоемкой операцией.
 
@@
 Известные детерминированные методы решения задач данного класса основаны, как правило, на подходах Branch-and-Bound \cite{Belotti} or Branch-and-Reduce \cite{Vigerske}.
-Также известен ряд метаэвристических и генетических алгоритмов, так или иначе основанных на идеях случайного поиска \cite{Deep,Schluter}.
+Для решения таких задач могут применяться также nature-inspired algorithms, так или иначе основанных на идеях случайного поиска \cite{Deep,Schluter}.
 
 %Ссылки на параллельные генетические алгоритмы - простота распараллеливания
-Вопросам распараллеливания генетических алгоритмов посвящены тысячи статей, т.к. данный класс алгоритмов допускает естественное распараллеливание.
+%Вопросам распараллеливания генетических алгоритмов посвящены тысячи статей, т.к. данный класс алгоритмов допускает естественное распараллеливание.
 
@@
 
-Отметим ключевые моменты, специфичные для построения детерминированных параллельных алгоритмов решения задач глобальной оптимизации.
+Отметим ключевые моменты, специфичные для построения параллельных алгоритмов решения mixed-integer global optimization problems.
 Организовать распараллеливание процесса поиска можно несколькими способами.
@@
 	\item
-	%локальное ускорение
+	учитывается информация о неоднородном поведении функций задачи в разных подобластях областях области поиска, что позволяет ускорить процесс поиска решения за счет совмещения процедур локального уточнения текущего лучшего решения и его глобального обновления;
 	\item 
@@
 
+Предположение о выполнении условия Липшица является типичным для многих подходов к построению параллельных алгоритмов глобальной оптмизации \cite{Evtushenko2009,Zilinskas2011}.
+При этом для решения многомерных задач оптимизации адаптируются алгоритмы решения одномерных задач, см., например, методы на основе search domain partitions \cite{Zilinskas2014,Sergeyev2017}.
+
 %Также будем предполагать, что функции задачи $g_i(y)$ определены и вычислимы только в тех точках обалсти $y \in D$, в которых выполняются ограничения
@@
 %\end{equation}
-
+В данной работе мы будем использовать подход к сведению многомерных задач к одномерным на основе space-filling curves. 
 Используя отображения типа Peano-Hilbert curve $y(x)$, отображающие отрезок $[0,1]$ на $N$-dimensional domain (\ref{D}),
@@
 
-Методы для построения чиcленных аппроксимаций Peano-Hilbert curve $y(x)$ (называемых \textit{evolvent}) рассмотрены в \cite{Strongin2013}.
+Методы для построения чиcленных аппроксимаций Peano-Hilbert curve $y(x)$ (называемых \textit{evolvent}) рассмотрены в \cite{Sergeyev2013}.
 These evolvents are fractals generated by an iterative process, that fill in the hypercube $D$ with accuracy $2^{-M}$, where integer $M>0$ is the evolvent construction parameter.
@@
 \begin{figure}
-\begin{minipage}{0.49\linewidth}
+\begin{minipage}{0.4\linewidth}
 \center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (a)}
 \end{minipage}
-\begin{minipage}{0.49\linewidth}
+\begin{minipage}{0.4\linewidth}
 \center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (b)}
@@
 
-Указанная схема редукции размерности сопоставляет многомерной задаче с липшицевой минимизируемой функцией и липшицевыми ограничениями одномерную задачу, в которой соответствующие функции удовлетворяют uniform H{\"o}lder condition (see \cite{Strongin2013}), i.e.,
+Указанная схема редукции размерности сопоставляет многомерной задаче с липшицевой минимизируемой функцией и липшицевыми ограничениями одномерную задачу, в которой соответствующие функции удовлетворяют uniform H{\"o}lder condition (see \cite{Sergeyev2013}), i.e.,
 \[
@@
 
-%эффективный алгоритм
+В работах проф.
+Стронгина был предложен эффективный параллельный алгоритм для решения constrained global optimization problems, основанный на индексной схеме учета ограничений.
+Приведем описание алгоритма в соответствии с \cite{Strongin2000,Strongin2013}.
 %исходная информация - параллельность
+
 Пусть в нашем распоряжении имеется $p$ вычислительных элементов (ядер или процессоров), которые будут использоваться для параллельного проведения $p$ испытаний одновременно.
 На первой итерации метода $p$ испытаний проводятся параллельно в произвольных различных точках $x^i\in(0,1)$, $1\leq i \leq p$.
-
 Пусть выполнено $n\geq 1$  итераций метода, в процессе которых были проведены испытания в $k=np$ точках $x^i, 1\leq i \leq k$.
@@
 
+Данный способ организации параллельных вычислений имеет следующее обоснование.
+Используемые в  характеристики интервалов (\ref{R}) могут рассматриваться как некоторые меры вероятности локализации в данных интервалах точки глобального минимума.
+Неравенства (\ref{Eq:23}) упорядочивают интервалы по их характеристикам, и испытания проводятся параллельно в первых $p$ интервалах, имеющих наибольшие вероятности.
+
 A detailed description of the algorithm convergence theory is presented in \cite{Strongin2000,Strongin2013}.
-
-%комментарий по поводу r и eps-резервов
+Роль параметров алгоритма ($r_\nu>1$ and $\epsilon_\nu>0, 1\leq\nu\leq m+1$) обсуждается в \cite{Strongin2020}.
+Изложенный алгоритм может быть также использован для решения задач многокритериальной оптимизации \cite{Gergel2020}.
 
@@
 В некоторых подобластях значения функций могут изменяться достаточно быстро (что будет соответствовать большим значениям константы Липшица в таких подобластях), в других подобластях значения могут изменяться более плавно.
-Как результат, построение единых оценок характеристик оптимизируемых функций для всей области поиска может не полностью соответствовать ее поведению в конкретной подобласти.
-
-Среди возможных подходов, учитывающих неоднородность поведения оптимизируемых функций в разных подобластях области поиска, можно выделить направление, связанное с адаптивным прогнозированием положения экстремума на основе накопленной поисковой информации в предположении слабой многоэкстремальности задачи.
+Как результат, построение единых оценок характеристик функции для всей области поиска может не полностью соответствовать ее поведению в конкретной подобласти.
 
@@
 
+Среди возможных подходов, учитывающих неоднородность поведения оптимизируемых функций в разных подобластях области поиска, можно выделить направление, связанное с адаптивным прогнозированием положения экстремума на основе накопленной поисковой информации в предположении слабой многоэкстремальности задачи.
 А именно, можно построить адаптивные правила прогнозирования, основанные на предположении о том, что местонахождение глобального оптимума более вероятно в окрестностях тех точек из ряда $x_i, 1 \leq i \leq k,$ которым соответствуют малые значения нормированных разностей $(z_i-z_\nu^*)/\mu_\nu$
@@
 
-Конкретным примером новых формул для вычисления характеристик $R'(i)$, включающих в себя адаптивый прогноз расположения оптимума, являются выражения
+Новые формулы для вычисления характеристик поисковых интервалов $R'(i)$, которые должны быть использованы на 5-м шаге алгоритма, могут быть представлены в виде
 \begin{eqnarray}\label{lR}
@@
 Управление параметром $q$, входящим в (\ref{lR}), позволяет учитывать различные предположения о характере экстремума.
-Например, одной из стратегий является чередование с заданной частотой больших и малых значений $q$, что соответствует смеси локального и глобального поиска, которую можно интерпретировать как сочетание процедуры локального уточнения текущего решения и процедуры глобального обновления этого решения.
+Например, одной из стратегий управления является чередование с заданной частотой больших и малых значений $q$, что соответствует смеси локального и глобального поиска, которую можно интерпретировать как сочетание процедуры локального уточнения текущего решения и процедуры глобального обновления этого решения.
 При этом в течение первых итераций рекомендуется использовать значение $q=1$, т.к.
 накопленная на них поисковая информация обеспечит адекватность формул (\ref{lR}) при последующем локальном уточнении.
-
-%Отметим, что на скорость сходимости алгоритмов глобальной оптимизации существенно влияет константа Липшица,
-%поэтому столь важным является вопрос ее корректной оценки.
-Заниженная оценка истинной константы Липшица L может привести к потере глобального решения.
-В то же время завышенная оценка константы L, не соответствующая истинному поведению функции, влечет за собой медленную сходимость алгоритма к точке глобального минимума.
-%Использование только глобальной информации о поведении целевой функции в ходе ее минимизации может значительно замедлить сходимость алгоритма к точке глобального минимума, т.к.
-могут существовать зоны области поиска, которые находятся далеко от точки глобального минимума, но локальные константы Липшица в которых существенно меньше глобальной.
-%
-%В работах [47, 140, 156] был предложен метод локальной настройки, преодолевающий указанные сложности и позволяющий адап-
-%тивно оценивать локальные константы Липшица на всей области поиска. Балансирование локальной и глобальной информации в про-
-%цессе поиска глобального минимума позволяет существенно ускорить работу алгоритма глобальной оптимизации.
-%
-%
-%Суть предложенного в подхода состоит в следующем.
-В алгоритме глобального поиска предлагается использовать не единую для всей области поиска оценку константы Липшица, а строить ее оценки для каждого поискового интервала в отдельности.
-Поскольку получение таких оценок может происходить с достаточно большими погрешностями, то при нахождении локальных (интервальных) оценок должна учитываться и общая (глобальная) оценка константы Липшица.
-%
-%Приведем алгоритмическое описание предлагаемого подхода.
-%Предположим, что параллельный алгоритм глобального поиска выполнил $k\geq$ итераций поиска и получены точки испытаний из (ссылка).
-%
-%Оценка константы Липшица   интервала (xi1,xi), 1<i<k, вычисляется в соот-ветствии со следующими выражениями:
 
@@
 шаг 7 алгоритма), которое может быть реализовано как на CPU (с использованием OpenMP и/или MPI), так и на GPU (с использованием CUDA).
+В данной работе для распараллеливания были задействованы только центральные процессоры.
+Распределение работ между узлами осуществлялось в соответствии с многоуровневой схемой декомпозиции параллельных вычислений, изложенной в \cite{Strongin2018,Barkalov2020}.
 
@@
 
-%\begin{figure}
-%\begin{minipage}{0.5\linewidth}
-%\center{\includegraphics[width=1.0\linewidth]{GKLS-4.png} \\ (a)}
-%\end{minipage}
-%\hfill
-%\begin{minipage}{0.5\linewidth}
-%\center{\includegraphics[width=1.0\linewidth]{GKLS-6.png} \\ (b)}
-%\end{minipage}
-%\caption{Solving a two-dimensional problem using the index algorithm}
-%\label{example}
-%\end{figure}
+\begin{figure}
+\begin{minipage}{0.48\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{GKLS-4.png} \\ (a)}
+\end{minipage}
+\hfill
+\begin{minipage}{0.48\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{GKLS-6.png} \\ (b)}
+\end{minipage}
+\caption{Solving a two-dimensional problem using the global search algorithm}
+\label{example}
+\end{figure}
 
@@
 
+\bibitem{Evtushenko2009} 
+Yu.~G.~Evtushenko, V.~U.~Malkova and A.~A.~Stanevichyus, \textit{Parallel global optimization of functions of several variables}, Computational Mathematics and Mathematical Physics \textbf{49}(2), 246--260 (2009).
+
+\bibitem{Zilinskas2011}
+R.~Paulavi\v{c}ius, J.~\v{Z}ilinskas and A.~Grothey, \textit{Parallel branch and bound for global optimization with combination of Lipschitz bounds}.
+Optimization Methods \& Software \textbf{26}(3), 487--498 (2011).
+
+\bibitem{Zilinskas2014}
+R.~Paulavi\v{c}ius and J.~\v{Z}ilinskas, \textit{Simplicial Global Optimization} (Springer Briefs in Optimization. Springer, 2014).
+
+\bibitem{Sergeyev2017}
+Y.~D.~Sergeyev and D.~E.~Kvasov, \textit{Deterministic Global Optimization. An Introduction to the Diagonal Approach}  (Springer Briefs in Optimization, Springer, 2017).
+
+%\bibitem{Evtushenko2013} Y.~G.~Evtushenko and M.~A.~Posypkin, \textit{A deterministic approach to global box-constrained optimization}, Optim. Lett. \textbf{7}(4), 819--829 (2013).
+
 \bibitem{Strongin2000}
@@
 
+\bibitem{Sergeyev2013}
+Ya.~D.~Sergeyev, R.~G.~Strongin and D.~Lera, \textit{Introduction to Global Optimization Exploiting Space-Filling Curves} (Springer Briefs in Optimization, Springer, 2013).
+
+\bibitem{Strongin2020}
+R.~Strongin, K.~Barkalov and S.~Bevzuk, \textit{Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints}, Soft Computing \textbf{24}(16), 11853--11865 (2020).
+
+\bibitem{Gergel2020}
+V.~Gergel, E.~Kozinov and K.~ Barkalov, \textit{Computationally efficient approach for solving lexicographic multicriteria optimization problems}, Optim. Lett. (2020). DOI: 10.1007/s11590-020-01668-y
+
+\bibitem{Strongin2018}
+R.~G.~Strongin, V.~P.~Gergel, K.~A.~Barkalov and A.~V.~Sysoyev, \textit{Generalized parallel computational schemes for time-consuming global optimization}, Lobachevskii Journal of Mathematics \textbf{39}(4), 576--586 (2018.)
+
+\bibitem{Barkalov2020}
+R.~G.~Strongin, V.~P.~Gergel and K.~A.~Barkalov, \textit{Adaptive global optimization based on a block-recursive dimensionality reduction scheme}, Automation and Remote Control \textbf{81}(8), 1475--1485 (2020).
+
 \bibitem{Floudas}
 C.~A.~Floudas and M.~P.~Pardalos,  \textit{Handbook of Test Problems in Local and Global Optimization}. (Springer, 1999).  %; DOI: 10.1007/978-1-4757-3040-1
+
+\bibitem{Gaviano}
+M.~Gaviano, D.~Lera, D.~E.~Kvasov and Ya.~D.~Sergeyev, \textit{Software for generation of classes of test functions with known local and global minima for global optimization}, ACM Trans.
+Math.
+Softw.
+\textbf{29}, 469--480 (2003).
 
@@
 %
-%\bibitem{Sergeyev1}
-%Y.~D.~Sergeyev and D.~E.~Kvasov, \textit{Deterministic Global Optimization. An Introduction to the Diagonal Approach}  (Springer Briefs in Optimization, Springer, 2017).
-%
-%\bibitem{Paulavicius}
-%R.~Paulavi\v{c}ius, J.~\v{Z}ilinskas, \textit{Simplicial Global Optimization} (Springer Briefs in Optimization. Springer, 2014).
 %
@@
 %G.~Luque and E.~Alba, \textit{Parallel genetic algorithms. Theory and real world applications} (Springer-Verlag, Berlin, 2011).
-%
-%\bibitem{Sergeyev2}
-%Ya.~D.~Sergeyev, R.~G.~Strongin, D.~Lera, \textit{Introduction to Global Optimization Exploiting Space-Filling Curves} (Springer Briefs in Optimization, Springer, 2013).
 %
@@
 %
-%\bibitem{Gaviano}
-%M.~Gaviano, D.~Lera, D.~E.~Kvasov and Ya.~D.~Sergeyev, \textit{Software for generation of classes of test functions with known local and global minima for global optimization}, ACM Trans.
-Math.
-Softw.
-\textbf{29}, 469--480 (2003).
-%
 %\bibitem{Modorskii}

### 2021-02-03 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
Some updates in Russian

@@
 Данный способ организации параллельных вычислений имеет следующее обоснование.
-Используемые в  характеристики интервалов (\ref{R}) могут рассматриваться как некоторые меры вероятности локализации в данных интервалах точки глобального минимума.
+Используемые в вычислительных правилах алгоритма характеристики интервалов (\ref{R}) могут рассматриваться как некоторые оценки вероятности обнаружения в данных интервалах точки глобального минимума.
 Неравенства (\ref{Eq:23}) упорядочивают интервалы по их характеристикам, и испытания проводятся параллельно в первых $p$ интервалах, имеющих наибольшие вероятности.
@@
 
+Детально проработанным направлением здесь является использование разных оценок константы Липшица для каждого поискового интервала в отдельности \cite{Sergeyev2003,Sergeyev2007}.
+Определенная сложность здесь состоит в том, что при построении локальных оценок константы Липшица неявно предполагается квадратичное поведение функции в малой окрестности исследуемой точки, что может негативно сказаться на работе метода вне этой окрестности.
+Одновременно с этим при вычислении локальных (интервальных) оценок должна учитываться и общая (глобальная) оценка константы Липшица, что приводит к использованию сложных компромиссных схем оценивания константы \cite{Sergeyev2020}.
+
 %Априорная информация, имеющаяся при решении прикладных многоэкстремальных задач, помимо факта ограниченности разделенных разностей функций задачи может содержать дополнительные характеристики экстремума, использование которых ускоряет процесс поиска.
@@
 
-Среди возможных подходов, учитывающих неоднородность поведения оптимизируемых функций в разных подобластях области поиска, можно выделить направление, связанное с адаптивным прогнозированием положения экстремума на основе накопленной поисковой информации в предположении слабой многоэкстремальности задачи.
-А именно, можно построить адаптивные правила прогнозирования, основанные на предположении о том, что местонахождение глобального оптимума более вероятно в окрестностях тех точек из ряда $x_i, 1 \leq i \leq k,$ которым соответствуют малые значения нормированных разностей $(z_i-z_\nu^*)/\mu_\nu$
-где $z_i$ из (\ref{trial_result}), $\mu_\nu$ из (\ref{Eq:20}), а $z_\nu^*$ из (\ref{Eq:21}) .
+Более удобным подходом, учитывающим неоднородность поведения функций задачи в разных подобластях области поиска,  является адаптивное прогнозирование положения оптимума на основе накопленной поисковой информации в предположении слабой многоэкстремальности задачи.
+Основы подхода изложены в \cite{Strongin2000}, некоторые результаты решения задач на параллельных вычислительных системах приведены в \cite{Barkalov2010}.
 
-Принятая нами гипотеза правдоподобна либо в случае, когда число осуществленных итераций достаточно велико (т.к.
+% здесь нужны ссылки
+В соответствии с данным подходом можно построить адаптивные правила прогнозирования, основанные на предположении о том, что местонахождение глобального оптимума более вероятно в окрестностях тех точек из ряда $x_i, 1 \leq i \leq k,$ которым соответствуют малые значения нормированных разностей $\omega_i=(z_i-z_\nu^*)/\mu_\nu$, где $\mu_\nu$ из (\ref{Eq:20}), а $z_\nu^*$ из (\ref{Eq:21}).
+
+Принятая гипотеза правдоподобна либо в случае, когда число осуществленных итераций достаточно велико (т.к.
 алгоритм при выполнении условий сходимости порождает последовательность испытаний, плотную лишь в точках глобального минимума), либо в случае, когда задача является унимодальной (т.е.
@@
 Новые формулы для вычисления характеристик поисковых интервалов $R'(i)$, которые должны быть использованы на 5-м шаге алгоритма, могут быть представлены в виде 
+\[
+R'(i) = h_i R(i),
+\]
+где $R(i)$ из (\ref{R}), а множитель $h_i$ вычисляется по правилам
 \begin{eqnarray}\label{lR}
-& R'(i) = \frac{R(i)}{\sqrt{(z_i-z_\nu^*)(z_{i-1}-z_\nu^*)}/\mu_\nu+1.5^{-q}}, \; \nu(x_{i-1})=\nu(x_{i}),\nonumber\\
-& R'(i) = \frac{R(i)}{(z_i-z_\nu^*)/\mu_\nu+1.5^{-q}}, \; \nu(x_{i-1})<\nu(x_{i}),\\
-& R'(i) = \frac{R(i)}{(z_{i-1}-z_\nu^*)/\mu_\nu+1.5^{-q}}, \; \nu(x_{i-1})>\nu(x_{i}),\nonumber
+& h_i = \left[\frac{\sqrt{(z_i-z_\nu^*)(z_{i-1}-z_\nu^*)}}{\mu_\nu}+1.5^{-q}\right]^{-1}, \; \nu(x_{i-1})=\nu(x_{i}),\nonumber\\
+& h_i = \left[\frac{z_i-z_\nu^*}{\mu_\nu}+1.5^{-q}\right]^{-1}, \; \nu(x_{i-1})<\nu(x_{i}),\\
+& h_i = \left[\frac{z_{i-1}-z_\nu^*}{\mu_\nu}+1.5^{-q}\right]^{-1}, \; \nu(x_{i-1})>\nu(x_{i}).\nonumber
 \end{eqnarray}
-где $R(i)$ из (\ref{R}) , а целочисленный параметр $q$ характеризует «степень локальности» алгоритма. При $q=1$ поиск имеет «глобальный», а при $q>>1$ -- «локальный» характер. 
-%Формулы (\ref{localR}) следует использовать в правилах алгоритма на шаге вместо
+Здесь целочисленный неотрицательный параметр $q$ характеризует «степень локальности» поиска: при $q=0$ поиск будет иметь ``глобальный'', а при $q>>1$ -- ``локальный'' характер. 
+В самом деле, множитель $h_i$ будет принимать наибольшее значение, равное $1.5^q$, лишь для тех интервалов, граничные точки которых соответствуют текущему минимуму, т.е.
+для которых выполняется равенство $\omega_i = 0$.
+Указанное значение множителя $h_i$ будет большим при больших значениях параметра $q$ (что увеличит значение характеристики данного интервала), и будет равно 1 при $q=0$ (что соответствует исходному значению характеристики).
 
@@
 Таким образом, мы сформировали Mixed-Integer Global Search Algorithm (MIGSA), основанный на сведении задачи mixed-integer non-convex optimization problem к одновременно решаемому набору непрерывных задач.
-Можно доказать, что условия сходимости данного алгоритма будут следовать из условий сходимости его прототипа (GSA).
 
@@
 Распределение работ между узлами осуществлялось в соответствии с многоуровневой схемой декомпозиции параллельных вычислений, изложенной в \cite{Strongin2018,Barkalov2020}.
-
-%здесь - сослаться на распараллеливание с использованием блочной многошаговой схемы.
 
@@
 
+\bibitem{Sergeyev2003}
+Y.~D.~Sergeyev, P.~Pugliese and D.~Famularo, \textit{Index information algorithm with local tuning for solving multidimensional global optimization problems with multiextremal constraints}, Math.
+Program.
+\textbf{96}(3), 489--512 (2003).
+2003.
+
+\bibitem{Sergeyev2007}
+Y.~D.~Sergeyev, D.~E.~Kvasov and F.~M.~H.~Khalaf, \textit{A one-dimensional local tuning algorithm for solving GO problems with partially defined constraints} Optimization Letters \textbf{1}(1), 85--99 (2007).
+
+\bibitem{Sergeyev2020}
+D.~E.~Kvasov, M.~S.~Mukhametzhanov, M.~C.~Nasso and Y.~D.~Sergeyev, \textit{On acceleration of derivative-free univariate Lipschitz global optimization methods}, Lecture Notes in Computer Science \textbf{11974}, 413--421 (2020).
+
+\bibitem{Barkalov2010}
+K.~Barkalov, V.~Ryabov, S.~Sidorov, \textit{Parallel scalable algorithms with mixed local-global strategy for global optimization problems}, Lecture Notes in Computer Science \textbf{6083}, 232–240 (2010).
+
 \bibitem{Strongin2018}

### 2021-02-04 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
Before translation of the first part

@@
 \begin{abstract} % You shouldn't use formulas and citations in the abstract.
-In this example, the article contains some required author
-information and examples of how to gain an article in the REV\TeX~4
-for \ljm. You shouldn't use formulas and citations in the abstract.
+%абстракта пока нет, подготовим позже
 \end{abstract}
@@
 В работе рассматриваются global optimization problems и параллельные методы их решения.
-Задачи глобальной оптимизации обладают значительной вычислительной трудоемкостью, т.к.
+Методы решения задач указанного класса  обладают значительной вычислительной трудоемкостью, т.к.
 глобальный оптимум является интегральной характеристикой решаемой задачи и требует исследования всей области поиска.
-Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (вообще говоря, неравномерного) в области параметров.
-Особый интерес представляют задачи, в которых часть параметров может принимать лишь целочисленные значения (mixed-integer global optimization problems), т.к.
-для них сложнее построить оценки оптимума по сравнению с непрерывными задачами.
+Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (как правило, неравномерного) в области изменения параметров.
+Особый интерес представляют задачи, в которых часть параметров может принимать лишь дискретные (например, целочисленные) значения (mixed-integer global optimization problems), т.к.
+для них сложнее построить оценки оптимума по сравнению с задачами, в которых все параметры являются непрерывными.
 
-Ситуация, когда некоторые из параметров характеризуются дискретностью (в частности, целочисленностью), является широко распространенной в прикладных оптимизационных задачах.
-При этом целочисленные параметры имеют, как правило, небольшое число значений и могут определять, например, марку используемого материала, вариант типовой компоновки деталей и т.п.
-При этом целевая функция и ограничения обычно заданы не явно, а в виде некоторого алгоритма вычисления их значений в точках области поиска, т.е.
-представлять собой “black-box”-функции.
-Процесс проверки ограничений и вычисления значения целевой функции в некоторой точке (называемый в дальнейшем trial в этой точке) требует проведения численного моделирования и является трудоемкой операцией.
+Ситуация, когда некоторые из параметров характеризуются дискретностью, является широко распространенной в прикладных оптимизационных задачах.
+При этом дискретные параметры имеют, как правило, небольшое число значений и могут определять, например, марку используемого материала, вариант типовой компоновки деталей и т.п.
+Одновременно с этим целевая функция и ограничения обычно заданы не явно, а в виде некоторого алгоритма вычисления их значений в точках области поиска, т.е.
+представляют собой функции вида “black-box”.
+Процесс проверки ограничений и вычисления значения целевой функции в некоторой точке (называемый в дальнейшем trial) требует проведения численного моделирования и является трудоемкой операцией.
 
@@
 Известные детерминированные методы решения задач данного класса основаны, как правило, на подходах Branch-and-Bound \cite{Belotti} or Branch-and-Reduce \cite{Vigerske}.
-Для решения таких задач могут применяться также nature-inspired algorithms, так или иначе основанных на идеях случайного поиска \cite{Deep,Schluter}.
-
-%Ссылки на параллельные генетические алгоритмы - простота распараллеливания
-%Вопросам распараллеливания генетических алгоритмов посвящены тысячи статей, т.к. данный класс алгоритмов допускает естественное распараллеливание.
-
-%Распараллеливание детерминированных алгоритмов - сложная задача
+Для решения таких задач могут применяться также nature-inspired algorithms \cite{Deep,Schluter}, которые так или иначе основаны на идеях случайного поиска.
 
@@
 Организовать распараллеливание процесса поиска можно несколькими способами.
-Во-первых, можно распараллеливать процесс проведения одного испытания, т.е.
+Во-первых, можно распараллеливать процесс проведения одного trial, т.е.
 вычисления функций задачи в одной точке области поиска.
@@
 
-В данной работе нами предложен новый parallel метод решения mixed-integer problems, основанный на information-statistical approach к решению задач глобальной оптимизации \cite{Strongin2000,Strongin2013}.
+В данной работе нами предложен новый параллельный метод решения mixed-integer problems, основанный на information-statistical approach к решению задач глобальной оптимизации \cite{Strongin2000,Strongin2013}.
 
@@
 В Section 4 предложен подход к обобщению параллельного индексного алгоритма с целью решения mixed-integer problems.
-Приведен способ, позволяющий свести решение mixed-integer problem к решению набора задач непрерывной оптимизации, которое может выполняться параллельно.
+Приведен способ, позволяющий свести решение mixed-integer problem к решению информационно-связанного множества задач непрерывной оптимизации, которое может выполняться параллельно.
 Section 5 содержит результаты численных экспериментов.
-Здесь производится сравнение последовательной версии алгоритма с известными аналогами, а также показана эффективность параллельной версии алгоритма с локальной настройкой при решении серии multiextremal mixed-integer problem.
+Здесь производится сравнение последовательной версии алгоритма с известными аналогами, а также показана эффективность распараллеливания алгоритма с локальной настройкой при решении серии multiextremal mixed-integer problem на суперкомпьютере.
 Section 6 concludes the paper.
@@
 
-Рассмотрим задачу глобальной оптимизации вида
+Рассмотрим задачу глобальной оптимизации с непрерывными параметрами вида
 \begin{eqnarray}\label{problem}
@@
 Предположение о выполнении условия Липшица является типичным для многих подходов к построению параллельных алгоритмов глобальной оптмизации \cite{Evtushenko2009,Zilinskas2011}.
-При этом для решения многомерных задач оптимизации адаптируются алгоритмы решения одномерных задач, см., например, методы на основе search domain partitions \cite{Zilinskas2014,Sergeyev2017}.
+При этом для решения многомерных задач оптимизации адаптируются алгоритмы решения одномерных задач, см., например, методы на основе разбиений области поиска \cite{Zilinskas2014,Sergeyev2017}.
 
-%Также будем предполагать, что функции задачи $g_i(y)$ определены и вычислимы только в тех точках обалсти $y \in D$, в которых выполняются ограничения
-%\begin{equation}\label{g_k}
-%g_k(y) \leq 0, \; 1 \leq k < i.
-%\end{equation}
 В данной работе мы будем использовать подход к сведению многомерных задач к одномерным на основе space-filling curves. 
@@
 
-В работах проф.
-Стронгина был предложен эффективный параллельный алгоритм для решения constrained global optimization problems, основанный на индексной схеме учета ограничений.
-Приведем описание алгоритма в соответствии с \cite{Strongin2000,Strongin2013}.
-%исходная информация - параллельность
+Рассмотрим параллельный алгоритм для решения constrained global optimization problems, предложенный в работах \cite{Strongin2000,Strongin2013} и основанный на индексной схеме учета ограничений.
 
-Пусть в нашем распоряжении имеется $p$ вычислительных элементов (ядер или процессоров), которые будут использоваться для параллельного проведения $p$ испытаний одновременно.
-На первой итерации метода $p$ испытаний проводятся параллельно в произвольных различных точках $x^i\in(0,1)$, $1\leq i \leq p$.
+При описании параллельного алгоритма мы будем предполагать, что в нашем распоряжении имеется $p$ вычислительных элементов (узлов, процессоров или ядер), которые будут использоваться для параллельного проведения $p$ trials одновременно.
+
+На первой итерации метода $p$ trials проводятся параллельно в произвольных различных точках $x^i\in(0,1)$, $1\leq i \leq p$.
 Пусть выполнено $n\geq 1$  итераций метода, в процессе которых были проведены испытания в $k=np$ точках $x^i, 1\leq i \leq k$.
@@
 
-%Проверить перевод!
 \begin{enumerate}
@@
 
-%комментарий про различные схемы распараллеливания - со ссылками на наши работы
-
 \section{Local tuning for global optimization algorithm}
@@
 
-%Априорная информация, имеющаяся при решении прикладных многоэкстремальных задач, помимо факта ограниченности разделенных разностей функций задачи может содержать дополнительные характеристики экстремума, использование которых ускоряет процесс поиска.
-К числу таких характеристик можно отнести некоторый априорный прогноз расположения оптимальной точки (основанный, например, на наличии прототипа оптимизируемого объекта).
-%Если же такой априорный прогноз отсутствует,
-
 Более удобным подходом, учитывающим неоднородность поведения функций задачи в разных подобластях области поиска,  является адаптивное прогнозирование положения оптимума на основе накопленной поисковой информации в предположении слабой многоэкстремальности задачи.
@@
 
-% здесь нужны ссылки
 В соответствии с данным подходом можно построить адаптивные правила прогнозирования, основанные на предположении о том, что местонахождение глобального оптимума более вероятно в окрестностях тех точек из ряда $x_i, 1 \leq i \leq k,$ которым соответствуют малые значения нормированных разностей $\omega_i=(z_i-z_\nu^*)/\mu_\nu$, где $\mu_\nu$ из (\ref{Eq:20}), а $z_\nu^*$ из (\ref{Eq:21}).
@@
 Здесь целочисленный неотрицательный параметр $q$ характеризует «степень локальности» поиска: при $q=0$ поиск будет иметь ``глобальный'', а при $q>>1$ -- ``локальный'' характер. 
-В самом деле, множитель $h_i$ будет принимать наибольшее значение, равное $1.5^q$, лишь для тех интервалов, граничные точки которых соответствуют текущему минимуму, т.е.
+В самом деле, множитель $h_i$ будет принимать наибольшее значение равное $1.5^q$ лишь для тех интервалов, граничные точки которых соответствуют текущему минимуму, т.е.
 для которых выполняется равенство $\omega_i = 0$.
@@
 Например, одной из стратегий управления является чередование с заданной частотой больших и малых значений $q$, что соответствует смеси локального и глобального поиска, которую можно интерпретировать как сочетание процедуры локального уточнения текущего решения и процедуры глобального обновления этого решения.
-При этом в течение первых итераций рекомендуется использовать значение $q=1$, т.к.
+При этом в течение первых итераций рекомендуется использовать значение $q=0$, т.к.
 накопленная на них поисковая информация обеспечит адекватность формул (\ref{lR}) при последующем локальном уточнении.
@@
 
-Рассмотрим теперь соответствие 
+Рассмотрим теперь отображение 
 \[
@@
 \]
-отображающее любую точку интервала [0,S] на область $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функции 
+ставящее в соответствие точке из интервала [0,S] точку в области $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функции 
 \[
@@
 
-Таким образом, мы сформировали Mixed-Integer Global Search Algorithm (MIGSA), основанный на сведении задачи mixed-integer non-convex optimization problem к одновременно решаемому набору непрерывных задач.
+Таким образом, мы сформировали Parallel Mixed-integer Global search Algorithm with Local tuning (PMGAL), основанный на сведении задачи mixed-integer non-convex optimization problem к одновременно решаемому набору непрерывных задач.
 
@@
 
+%Ниже пока переводить не надо
+
 Первая серия экспериментов была проведена с использованием последовательной версии предложенного алгоритма с целью его сравнения с известными методами аналогичного назначения.
 
-Предложенный в данной работе алгоритм MIGSA сравнивался с genetic algorithm for solving mixed-integer optimization problems, реализованным в Matlab Global Optimization Toolbox.
+Предложенный в данной работе алгоритм PMGAL (в его последовательном варианте) сравнивался с genetic algorithm for solving mixed-integer optimization problems, реализованным в Matlab Global Optimization Toolbox.
 В Table \ref{tab:1} приведено число испытаний, потребовавшееся для решения известных тестовых mixed-integer problems данными методами.
@@
 Все вычислительные эксперименты были проведены на компьютере с процессором Intel Core i5-7300 2.5 GHz и 8 Gb RAM.
-Результаты экспериментов показывают превосходство метода MIGSA как по числу итераций, так и по времени работы.
+Результаты экспериментов показывают превосходство метода PMGAL как по числу итераций, так и по времени работы.
 
 \begin{table}
-	\caption{Сравнение эффективности методов MIGSA и GA}
+	\caption{Сравнение эффективности методов PMGAL и GA}
 	\label{tab:1}
@@
 	\[
-			\mu(y,u) = 2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} \right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right],
+			\phi (y,u) = 2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} \right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right],
 	\]
@@
 	\[
-	C = 4 - \min_{y,u} \left\{ \mu(y,u) \right\}, \; y\in D, u \in U.
+	C = 4 - \min_{y,u} \left\{ \phi(y,u) \right\}, \; y\in D, u \in U.
 	\]
-	Минимум вогнутой функции $-\mu(y,u)$ достигается в одной из угловых точек области поиска, поэтому при размерности задачи порядка 10 его вычисление может быть выполнено методом перебора.
+	Минимум вогнутой функции $-\phi(y,u)$ достигается в одной из угловых точек области поиска, поэтому при размерности задачи порядка 10 его вычисление может быть выполнено методом перебора.
 	\item Формируется многоэкстремальная mixed-integer function 
 	\[
-	\varphi(y,u) = \left(\psi(y) + \sum_{j=1}^M{u_j}\right)\left(C - \mu(y,u)\right).
+	\varphi(y,u) = \left(\psi(y) + \sum_{j=1}^M{u_j}\right)\left(C - \phi(y,u)\right).
 	\]

### 2021-02-08 2021/PСT_GergelKozinov/2021_PCT.tex
1 \omega -> \Omega

@@
 \begin{equation}\label{eq:13}
- \omega_k=\{ y^i,u^i,z^i=F(\alpha,y^i,u^i ): 1\leq i\leq k \}
+ \Omega_k=\{ y^i,u^i,z^i=F(\alpha,y^i,u^i ): 1\leq i\leq k \}
 \end{equation}
@@
 \begin{equation}\label{eq:14}
-	\theta =\Theta(\omega_k ),1 \leq \theta \leq l, 
+	\theta =\Theta(\Omega_k ),1 \leq \theta \leq l, 
 \end{equation}
@@
 
-\textit{Rule 1.} Apply the decision rule $\Theta$ from (\ref{eq:14}) to available search information $\omega_k$ from (\ref{eq:13}),  and determine the value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.
+\textit{Rule 1.} Apply the decision rule $\Theta$ from (\ref{eq:14}) to available search information $\Omega_k$ from (\ref{eq:13}),  and determine the value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.
 
@@
 \textit{Rule 3.} Check the stopping condition for the computation.
-If the required accuracy of the global search is not achieved, it is necessary to supplement the search information $\omega_k$ from (\ref{eq:13}) with the results of the current iteration and continue calculations starting from Rule 1.
+If the required accuracy of the global search is not achieved, it is necessary to supplement the search information $\Omega_k$ from (\ref{eq:13}) with the results of the current iteration and continue calculations starting from Rule 1.
 
@@
 The problem of increasing computational complexity was even called the ``dimensionality curse''.
-In particular, when applying the computational scheme discussed in subsection \ref{subsec:31}, it is necessary to accumulate and analyze multidimensional search information $\omega_k$ from (\ref{eq:13}).
+In particular, when applying the computational scheme discussed in subsection \ref{subsec:31}, it is necessary to accumulate and analyze multidimensional search information $\Omega_k$ from (\ref{eq:13}).
 This computational complexity can be significantly reduced by reducing the optimization problems to be solved by using Peano \textit{curves} or \textit{evolvents} $y(x)$, which unmbiguously and continuously map a segment [0,1] on an $n$-dimensional hyper-cube $D$ -- see, for example, \cite{c24,c25}.
@@
 
-As a result of dimensionality reduction, the search information $\omega_k$ from (\ref{eq:13}) obtained in the calculation can be presented in the form 
+As a result of dimensionality reduction, the search information $\Omega_k$ from (\ref{eq:13}) obtained in the calculation can be presented in the form 
 \begin{equation}\label{eq:17}

### 2021-02-11 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
1. Часть перевода motor 2021

@@
 %Пакеты для отображения русского шрифта
-\usepackage[utf8]{inputenc}
-\usepackage[russian]{babel}
+%\usepackage[utf8]{inputenc}
+%\usepackage[russian]{babel}
 %Пакеты для гиперссылок
@@
 %
-\title{Global search algorithm with local tuning for solving mixed-integer global optimization problems
+\title{An Approach for Simultaneous Finding of Multiple Effective Decisions in Multi-objective Optimization Problems
 %with non-convex constraints 
-\thanks{This study was supported by the .}}
+\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.0729-2020-0055, and by the Research and Education Mathematical Center, project no.
+075-02-2020-1483/1.}}
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Finding of Multiple Effective Decisions in MOO Problems}
 % If the paper title is too long for the running head, you can set
@@
 \author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
-Ilya Lebedev\orcidID{0000-0002-8736-0652}
+Victor Gergel\orcidID{0000-0002-4013-2329} \and
+Vladimir Grishagin \and
+Evgeniy Kozinov\orcidID{0000-0001-6776-0096} 
 }
 %
-\authorrunning{K. Barkalov, I. Lebedev}
+\authorrunning{K. Barkalov, V. Gergel, V. Grishagin, E. Kozinov}
 % First names are abbreviated in the running head.
@@
 \institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
-\email{\{konstantin.barkalov,ilya.lebedev\}@itmm.unn.ru}}
+\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru,\{gergel,vagris\}@unn.ru}
+}
 %
@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-150--250 words.
+This paper considers computationally intensive multicriteria optimization problems which require computing multiple Pareto-optimal decisions.
+It is also assumed that efficiency criteria may be multiextremal, and the cost of calculating function values may be quite high.
+The proposed approach is based on the reduction of multicriteria optimization problems to one-dimensional global optimization problems that can be solved using effective statistical information algorithms of global search.
+One key innovation of the developed approach consists in the possibility of solving several global optimization problems simultaneously, which allows multiple Pareto-optimal decisions to be obtained.
+Besides, such approach provides for reuse of the computed search information, which considerably reduces computational effort for solving multicriteria optimization problems.
+Computational experiments confirm the potential of the proposed approach.
 
-\keywords{First keyword  \and Second keyword \and Another keyword.}
+\keywords{Multi-objective optimization \and Multiple global optimization \and Dimensionality reduction \and Optimization method \and Search information \and Computational complexity.}
 \end{abstract}
@@
 %
-\section{First Section}
-\subsection{A Subsection Sample}
-Что-то по русски
-Please note that the first paragraph of a section or subsection is
-not indented. The first paragraph that follows a table, figure,
-equation etc. does not need an indent, either.
+\section{Introduction} \label{sec:01}
 
-Subsequent paragraphs, however, are indented.
+Choosing the best options (decision making) in situations with many different alternatives is a problem that occurs in almost every domain of human activity.
+In the simplest cases, decision making problems can be viewed as optimization problems of various kinds such as convex programming, discrete optimization, nonlinear programming, etc.
+In more complex situations, the objective functions that determine the effectiveness of decisions can be multiextremal, and decision making will require solving global optimization problems.
+In the most general case, however, the effectiveness of decisions may be determined by several objective criteria, which necessitates solving multi-objective optimization (MOO) problems.
+Thus, MOO problems are the most common decision making settings that one has to solve in many scientific and technical applications.
+A large number of approaches and methods have been developed to solve such problems and they have been used to solve many decision making problems in various fields of practical applications -- see e.g.
+\cite{c1,c2,c3,c4,c5,c6,c7,c8,c9}.
 
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
+Having several objective functions in MOO problems that determine the effectiveness of the decisions to be made results in most cases in situations where no option can be found in the whole set of alternatives that would be the best one with respect to all efficiency criteria at the same time.
+In such situations, the solution of MOO problems usually consists in finding non-dominated (or efficient) decisions in which values cannot be improved simultaneously across all objective functions.
+The set of all efficient decisions is usually referred to as the Pareto set.
 
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
+Finding the whole Pareto set when solving MOO problems can be computationally intensive (especially when objective functions are multiextremal).
+Moreover, having the whole Pareto set may be redundant for the decision maker (DM) due to the complexity of analyzing a large number of possible alternatives.
+As a result, when solving MOO problems in practice, it may be sufficient to find one or several efficient decisions taking into account the DM's perceptions of the required optimality of the options to be chosen.
 
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
+One of the approaches most commonly used in the search for efficient solutions is scalarization of the vector efficiency criterion, when a MOO problem is reduced to solving one or several scalar (in the general case, global) optimization problems in which the objective function is the only function.
+The family of optimization problems generated in such approach is further referred to as multiple global optimization problem (MGO).
+Within this approach, we can distinguish lexicographic optimization methods where objective functions are ordered by importance, thus  allowing the optimization of the functions to be carried out sequentially as their importance decreases [10].
+Possible scalarization methods also include various methods of the efficiency criteria convolution such as the weighted sum method, the compromise programming method, the reference point method, the weighted min-max method, etc.
+-- see, for example, \cite{c2,c11,c12}.
 
-\noindent Displayed equations are centered and set on a separate
-line.
+One common property of methods based on scalarization of the vector efficiency criterion is the existence of some scalarization coefficients that can be varied to obtain different solutions from the Pareto set.
+Thus, the scalarization coefficients may be interpreted as measures of the importance of the efficiency criteria determined by DM according to his/her perception of the required optimality of the decisions to be made.
+As a result, the general scheme for solving the MOO problem can be represented as a sequence of steps; at each step, DM sets the necessary scalarization coefficients, then the resulting scalar optimization problem is solved, after that DM analyzes the efficient decision found and, if necessary, the above steps are repeated.
+
+The general scheme discussed above can be extended by the possibility of selecting not one but several different scalarization coefficient options at each step.
+With this possibility, the task of coefficient assignment becomes less complex for DM.
+Solving several generated scalar optimization problems simultaneously allows one to get efficient decision estimates at the very early stages of computations thus making it possible to change dynamically (in the process of computations) the set of problems being solved: to stop solving obviously unproductive (from DM's point of view) ones or to add new optimization problems.
+
+It is also important to note that by solving simultaneously a large number of scalar optimization problems thus generated it is possible to significantly decrease computational complexity of each separate problem.
+This effect is achived due to the fact that all such scalar problems  are based on the same MOO problem and, consequently, all computed values of the efficiency criteria of the MOO problem can be reduced to values of any scalar problem being solved simultaneously without any time consuming calculations.
+In such cases, all the search information obtained in solving any single scalar problem can be used for solving all other scalar problems of the same group.
+
+The structure of the paper is as follows.
+In Section \ref{sec:02}, we give the statement of the multicriteria optimization problem.
+Section \ref{sec:03} presents a scheme for reduction of multicriteria optimization problems to one-dimensional global optimization problems as well as some methods for solving such problems.
+Section \ref{sec:04} considers the proposed approach for simultaneous solution of several global optimization problems yielding several Pareto-optimal decisions at a time.
+Section \ref{sec:05} contains the results of numerical experiments confirming the effectiveness of the proposed approach.
+In conclusion, the results obtained are discussed and possible main lines of action for further research are outlined.
+
+\section{Problem statement} \label{sec:02}
+
+In the most general form, the multi-objective optimization (MOO) problem can be formulated as follows
 \begin{equation}
-x + y = z
+\label{eq:01}
+f(y) = (f_1(y), f_2(y), \dots, f_s(y)) \to \min,  y\in D,
 \end{equation}
-Please try to avoid rasterized images for line-art diagrams and schemas. Whenever possible, use vector graphics instead (see Fig.~\ref{fig1}).
+where $f(y) = (f_1(y), f_2(y), \dots, f_s(y))$ are objective functions (efficiency criteria), $y = (y_1, y_2, \dots , y_N)$ is the vector of varied parameters, and a $N$ is the dimensionality of the multi-objective optimization problem to be solved.
+The set of possible parameter values (search domain) $D$ is usually an $N$-dimensional hypercube
+\begin{equation}
+\label{eq:02}
+D  = \{ y\in R^N: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}
+\end{equation}
+for given boundary vectors $a$ and $b$.
 
-%\begin{figure}
-%\includegraphics[width=\textwidth]{fig1.eps}
-%\caption{A figure caption is always placed below the illustration.
-%Please note that short captions are centered, while long ones are
-%justified by the macro package automatically.} \label{fig1}
-%\end{figure}
+Without loss of generality, it is assumed that objective functions should be minimized to improve the decision efficiency $y \in D$.
+It is also assumed that $f_i(y)$, $1 \leq i \leq s$, are multiextremal and have the form of time-consuming ``black-box'' computational procedures.
+It is also assumed that the objective functions.
+It is also assumed that the objective functions $f_i(y)$, $1 \leq i \leq s$, satisfy the Lipschitz condition
+\begin{equation}
+\label{eq:03}
+|f_i(y') - f_i(y'')| \leq L_i\|y' - y''\|, y', y'' \in D, 1 \leq i \leq s,
+\end{equation}
+where $L_i$, $1 \leq i \leq s$, are the Lipschitz constants and $\|*\|$ denotes the Euclidean norm in $R^N$.
+Condition (\ref{eq:03}) means that for small variations of the parameter $y \in D$, the corresponding changes in the values of the functions $f_i(y)$, $1 \leq i \leq s$, are bounded.
 
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-\cite{Strongin2000}.
+\section{Reducing the Problem of Multi-objective Optimization to the Problems of One-dimensional Global Optimization} \label{sec:03}
+
+As mentioned earlier, in the framework of the proposed approach the solution of MOO problems is reduced to solving one or several scalar global optimization problems in which a single efficiency criterion is generated using some scalarization methods of multiple objective functions.
+Below, we consider a general scheme for such reduction of the MOO problem and present a method for solving the generated global optimization problems.
+
+\textbf{1.
+Scalarization of multiple objective functions.} In the most general form, the global optimization problem generated by the scalarization of multiple objective functions of the MOO problem can be represented in the form
+\begin{equation}
+\label{eq:04}
+\min {\varphi(y)} = F(\lambda, y), y \in D,
+\end{equation}
+where $F$ is a scalar multiextremal function generated as a result of scalarization of objective functions $f_i$, $1 \leq i \leq s$, $\lambda$ is the vector of parameters of the applied convolution of functions, and $D$ is the search domain from (\ref{eq:01}).
+In the proposed approach, we use for scalarization the compromise programming method \cite{c2,c11}, where solution of the MOO problem consists in finding the efficient decision corresponding most closely to the optimality measures of the specified reference decision $y^0 \in D$.
+In this case, a possible variant of convolution of objective functions $f_i$, $1 \leq i \leq s$, can have the form
+\begin{equation}
+\label{eq:05}
+\min F(\lambda,y)=1/s \sum_{i=1}^s {\lambda_i (f_i (y)-f_i (y^0 ))^2}, y \in D,
+\end{equation}
+where $F(\lambda,y)$ is the standard deviation of the values of the objective functions $f_i$, $1 \leq i \leq s$, for the decision $y \in D$ and for the specified reference decision $y^0 \in D$, while the coefficients $0 \leq \lambda_i \leq 1$, $1 \leq i < s$, are measures of the significance of the approximation accuracy for each variable $y_i$, $1 \leq i \leq N$, separately  without loss of generality, we can assume that the area of possible values of the coefficients $\lambda$ is a set
+\begin{equation}
+\label{eq:06}
+\lambda=(\lambda_1,\lambda_2, \dots, \lambda_s)\in \Lambda \subset R^s: \sum_{i=1}^{s}\lambda_i=1, \lambda_i \geq 0, 1 \leq i \leq s.
+\end{equation}
+
+The reference decision $y^0 \in D$ in (\ref{eq:05}) can be known \textit{a priori} or determined on the basis of some known prototype.
+In many cases, an abstract ideal decision $y^0 \in D$ is used as the reference decision $y^*\in D$ in which the objective functions $f_i$, $1 \leq i \leq s$, assume minimal possible values, i.e.
+\begin{equation}
+\label{eq:07}
+f_i^* =f_i (y^*)=\min_{y \in D} f_i (y), 1 \leq i \leq s.
+\end{equation}
+
+Note that by virtue of (\ref{eq:03}) the scalar function $F(\lambda,y)$ from (\ref{eq:05}) also satisfies the Lipschitz condition with some constant $L$, i.e.
+\begin{equation}
+\label{eq:08}
+|F(\lambda, y') - F(\lambda, y'')| \leq L\|y' - y''\|.
+\end{equation}
+
+\textbf{2.
+Dimensionality reduction.} As already mentioned, problem (\ref{eq:04}) is a multidimensional global optimization problem.
+Problems of this kind are computationally complex and are known to be subject to the ``curse of dimensionality'' -- computational complexity increases exponentially with increasing dimensionality of the optimization problem being solved \cite{c13,c14,c15,c16,c17,c18,c19,c20}.
+Nevertheless, the computational complexity of global optimization algorithms can be significantly reduced by dimensionality reduction based on the use of a Peano space-filling curve (or evolvent) $y(x)$ that uniquely and continuously maps the segment $[0,1]$ on an $N$-dimensional hypercube $D$ -- see, for example, \cite{c15,c21}.
+As a result of such reduction, multidimensional global optimization problems (\ref{eq:04}) are reduced to one-dimensional problems
+\begin{equation}
+\label{eq:09}
+\min \varphi (y(x))= \min \varphi(y), x \in [0,1], y \in D.
+\end{equation}
+
+The resulting one-dimensional functions $\varphi(y(x))$ satisfy the uniform H{\" o}lder condition, i.e.
+\begin{equation}
+\label{eq:10}
+|\varphi (y(x'))- \varphi (y(x''))| \leq H |x'-x''|^{1/N},
+\end{equation}
+
+where the constant $H$ is defined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:08}), and $N$ is the dimensionality of the optimization problem (\ref{eq:01}).
+
+As a result of dimensionality reduction, the computational scheme for solving problem (\ref{eq:09}) consists of the following (see Fig. \ref{fig:1}):
+
+\begin{itemize}
+	\item The optimization algorithm minimizes the reduced one-dimensional function $\varphi(y(x))$ from (\ref{eq:09}),
+	\item After determining the point $x \in [0,1]$ of the next iteration of the one-dimensional global search, the multi-dimensional image $y \in D$ is computed for the mapping $y(x)$,
+	\item At the multi-dimensional point $y \in D$, the value of the original multi-dimensional function $\varphi(y)$ is computed from (\ref{eq:04}),
+	\item The computed value $z = \varphi(y)$ is then used as the value of the reduced one-dimensional function $\varphi(y(x))$ at  the point $x\in[0,1]$.
+\end{itemize}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{fig1}
+  \caption{General scheme for solving the problem (\ref{eq:09})}
+  \label{fig:1}
+\end{figure}
+
+\textbf{3.
+Solving the one-dimensional reduced optimization problem.} Using the dimensionality reduction results in one more additional advantage of the proposed approach: many well-known one-dimensional global search algorithms (possibly, after some additional generalisation) can be used to solve the initial multidimensional MOO problem from (\ref{eq:01}) -- see e.g.
+\cite{c22,c23,c24,c25,c26,c27,c28,c29}.
+At the same time, it should be noted that most works where dimensionality reduction is a key feature for solving multiextremal problems like (\ref{eq:04}) rely on the statistical information theory of global search \cite{c15}.
+This theory has provided the a basis for developing a large number of effective methods for multi-extremal optimization -- see, e.g., \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
+
+Within the framework of statistical information theory, a general computational scheme of global optimization algorithms was proposed which in brief is as follows \cite{c15,c30,c35}.
+
+Let $k$, $k \geq 2$, global search iterations aimed to minimize the function $\varphi(y(x))$ from (\ref{eq:09}) were completed.
+Then, to perform adaptive choice of the points of next iterations, the optimization algorithm estimates the possibility that the global minimum is located  in the intervals, into which the initial segment $[0,1]$ is divided by the points of earlier global search  iterations
+\begin{equation}
+\label{eq:11}
+x_1<x_2< \dots <x_k.
+\end{equation}
+
+This estimate is determined by means of characteristics of intervals $R(i)$, $1 \leq i<k$, whose values should be proportional to the degree of possibility that the global minimum is located  in these intervals.
+The type of these characteristics depends on the global optimization algorithm used -- thus, for example, for an algorithm which constructs a uniform dense grid in the global search area, the characteristic is simply the interval length
+\begin{equation}
+\label{eq:12}
+R(i)=(x_i-x_{i-1}), 1 \leq i<k.
+\end{equation}
+
+For the algorithms proposed in \cite{c22,c23}, when dimensionality reduction is applied, the characteristic is an estimate of the minimum possible value of the function $\varphi(y(x))$ to be minimized on the interval $(x_{i-1},x_i )$, $1 \leq i < k$, i.e.
+\begin{equation}
+\label{eq:13}
+R(i) = 0.5 H \varrho_i - 0.5 (z_{i-1} + z_i ), \varrho_i = (x_i-x_{i-1} ) ^ {1⁄N}, 1 \leq i < k,
+\end{equation}
+where $H$ is the H{\" o}lder constant from (\ref{eq:10}) for the reduced global optimization problem being solved (\ref{eq:09}), $z_i = \varphi(y(z_i))$, $1 \leq i < k$ and $N$ is the dimensionality of the problem from (\ref{eq:01})).
+For the global search algorithm (GSA) \cite{c15,c24} developed in the framework of the statistical information approach, the characteristic is
+\begin{equation}
+\label{eq:14}
+R(i)=m\varrho_i+\frac{z_i-z_{i-1}^2}{m\varrho_i}-2(z_{i-1}+z_i ),\varrho_i=(x_i-x_{i-1})^{1/N}, 1 \leq i < k,
+\end{equation}
+where $m$ is a numerical estimate of the H{\" o}lder constant derived from available search information
+\begin{equation}
+\label{eq:15}
+m= r M, M = \max |z_i-z_{i-1} | ⁄ \varrho_i, 1 \leq i<k
+\end{equation}
+ ($r>1$ is a parameter of the GSA algorithm).
+
+The presence of interval characteristics makes it possible to present the procedure of global search iteration as the following sequence of steps \cite{c15}.
+
+\textit{Step 1.} Calculate characteristics of the  intervals $R(i)$, $1 \leq i < k$, and determine the interval with the maximum characteristic
+\begin{equation}
+\label{eq:16}
+R(t) = \max R(i), 1\leq i < k.
+\end{equation}
+
+\textit{Step 2.} Select the next iteration point in the interval with the maximum characteristic
+\begin{equation}
+\label{eq:17}
+x^{k+1}=X(x_{t-1}, x_t),
+\end{equation}
+calculate the value $z^{k+1}$ of the function to be minimized at this point (the procedure for calculating the function value will be further referred to as a trial).
+
+\textit{Step 3.} Check the stopping condition
+\begin{equation}
+\label{eq:18}
+\varrho_t \leq \varepsilon,
+\end{equation}
+where $\varrho_t = (x_t-x_{t-1})^{1⁄N}$, $t$ from (\ref{eq:16}) and $\varepsilon > 0$ is the specified accuracy of the solution.
+If the stopping condition (\ref{eq:18}) is reached, then the solving of the optimization problem is stopped, otherwise $k=k+1$ is assumed and the next global search iteration begins.
+
+After completing the calculations, the lowest computed value of the function being minimized can be taken as the global minimum estimate
+\begin{equation}
+\label{eq:19}
+z_k^* = \min z_i, 1 \leq i<k.
+\end{equation}
+
+It should be noted again that the computational scheme discussed above is quite general.
+Many global search algorithms can be represented within this scheme, as evidenced, in particular, by examples (\ref{eq:12})-(\ref{eq:14}) and other characteristic algorithms -- see, for example, \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
+
+Convergence conditions for characteristic algorithms depend on the properties of interval characteristics used.
+One of sufficient conditions for convergence of algorithms is, for example, the requirement that the characteristic of the interval containing the global minimum point should take on the maximum value at step 1 of the characteristic scheme during the global search iterations.
+This condition is satisfied, for example, for the multidimensional generalized algorithms proposed in \cite{c22,c23} when the H{\"o}lder constant from (\ref{eq:10}) is specified exactly.
+For GSA, a sufficient condition for convergence is the relation \cite{c15}
+\begin{equation}
+\label{eq:20}
+m \geq 2^{3-1⁄N} L\sqrt{N+3},
+\end{equation}
+which must be fulfilled starting from some iteration $k>1$ of the global search ($L$ is the Lipschitz constant from (\ref{eq:08})).
+Moreover, if condition (\ref{eq:20}) is satisfied, only the points of the global minimum of the function $\varphi(y)$ from (\ref{eq:04}) will be the limit points of the trial sequence $\{y^k=y(x^k)\}$ generated by the GSA algorithm.
+
+\section{An Approach for Simultaneous Finding of Multiple Effective Solutions in Multi-objective Optimization Problems} \label{sec:04}
+
+\textbf{1.
+The need to solve multiple global optimization problems.} As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different effective decisions due to possible changes in the optimality requirements.
+Obtaining different  effective decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multi-extremal functions $F(\lambda, y)$ from (\ref{eq:04}).
+The resulting different functions $F(\lambda, y)$ for different values of the coefficients $\lambda \in \Lambda$ can be optimized sequentially.
+This deermines a multi-stage scheme for solving the MOO problem, when at each stage DM specifies the necessary scalarization coefficients, then the resulting scalar optimization problem is solved, after which DM analyzes the effective solution found.
+
+The general scheme considered above can be extended by the option of selecting not one but several different scalarization coefficients at each stage.
+Having such an option makes it easier for DM  to specify importance coefficients $\lambda \in \Lambda$ of the objective functions $f_i(y)$, $1 \leq i \leq s$.
+By solving simultaneously several generated scalar optimization problems, one can  obtain the estimates of efficient decisions at the earliest stages of computing, which makes it possible to dynamically (in the process of computation) change the set of problems being solved -- to stop solving those that obviously have no prospect of success (from DM's point of view) or to add new optimization problems.
+
+Such generalization of the process of solving a MOO problem means that at each current moment of calculations there is a set of functions being optimized simultaneously having the form (\ref{eq:04})
+\begin{equation}
+\label{eq:21}
+\Phi_p (y)=\{ F(\lambda_1,y), F(\lambda_2,y),\dots,F(\lambda_p,y)  \},  \lambda_i \in \Delta, 1 \leq i \leq p
+\end{equation}
+which can be changed dynamically in the course of the calculations by adding new or removing existing optimization functions $F(\lambda, y)$ from (\ref{eq:04}).
+
+\textbf{2.
+Step-by-step solution of a set of global optimization problems.} As shown in Section \ref{sec:03}, the statistical information multiextremal optimization algorithms used in the proposed approach determine the points of consecutive iterations of the global search taking into account the search information
+\begin{equation}
+\label{eq:22}
+A_k=\{(x_i,z_i,f_i )^T:1 \leq i \leq k\},
+\end{equation}
+obtained in the calculations (see (\ref{eq:14})-(\ref{eq:17})).
+In (\ref{eq:22}) $x_i$, $1 \leq i \leq k$,  are the reduced points of performed global search iterations ordered in ascending order of coordinates,  $z_i$, $f_i$, $1 \leq i \leq k$, are values of the scalar function $F(\lambda, y)$, from (\ref{eq:04}) and objective functions $f_i(y)$, $1 \leq i \leq s$, from (\ref{eq:01}) of the current optimization problem to be solved at the points $x_i=y(x_i)$, $1 \leq i \leq k$.
+By using search information $A_k$ from (\ref{eq:22}) when choosing next search iterations it is possible to solve global optimization problems more efficiently and to provide convergence of algorithms only to the global minima of multiextremal functions being minimized.
+
+It is important to note that since the set $\Phi_t(y)$ of functions being optimized  simultaneously is generated from the same MOO problem from (\ref{eq:01}), the existence of the set $A_k$ from (\ref{eq:12}) allows us to adjust the results of all previously performed calculations of the values of the objective functions $f_i(y)$, $1 \leq i \leq s$, to the values of the next optimized function $F(\lambda, y)$ from (\ref{eq:04}) without repeating any time consuming calculations of values, i.e.
+\begin{equation}
+\label{eq:23}
+(x_i,f_i ) \to z_i=F(\lambda, y(x_i)) ,1 \leq i \leq k.
+\end{equation}
+
+Thus, all the search information $A_k$ from (\ref{eq:22}), recalculated according to (\ref{eq:23}), can be reused to continue solving the problems of the set $\Phi_t(y)$.
+Such a possibility provides a significant reduction in computations up to performing only a limited set of global search iterations (see the results of computational experiments in Section \ref{sec:03}).
+
+     This type of information connectivity of functions in the set $\Phi_t(y)$ from (\ref{eq:21}) makes it possible to generalize the computational scheme (\ref{eq:16})-(\ref{eq:18}) for solving a single global optimization problem for the case of optimizing the functions of the set $\Phi_t(y)$ from (\ref{eq:21}) by adding a preliminary step of the search information  transformation.
+
+\textit{Step 0.} Adjust the state of search information $A_k$ from (\ref{eq:22}) to the values of the function $F(\lambda, y)$ from the set $\Phi_t(y)$ according to rule (\ref{eq:23}).
+
+The GSA algorithm applied to optimize the functions of the set $\Phi_t(y)$ from (\ref{eq:21}) and using search information $A_k$ will be further referred to as the Multiple Global Search Algorithm (MGSA).
+
+\textbf{3.
+Simultaneous solution of a set of global optimization problems.} Information connectivity makes it possible to propose a more general scheme for simultaneous optimization of all the functions of the set $\Phi_t(y)$ from (\ref{eq:21}).
+In this case, search information $A_k$ from (\ref{eq:22}) will contain the computed values of all simultaneously optimized functions $F(\lambda_i, y)$, $1 \leq i \leq p$, i.e.
+\begin{equation}
+\label{eq:24}
+A_k=\{(x_i,\overrightarrow{z_i},f_i )^T : 1 \leq i \leq k\}
+\end{equation}
+where the values of $\overrightarrow{z_i}$, $1 \leq i \leq k$, represent vectors 
+\begin{equation}
+\label{eq:25}
+\overrightarrow{z_i}=( z_i (1),z_i (2), \dots ,z_i (p)),z_i (j)=F(\lambda_j,y(x_i)), 1\leq i \leq k,1\leq j\leq p.
+\end{equation}
+
+Accordingly, for each interval $(x_{i-1},x_i)$, $1 \leq i < k$, into which the segment [0,1] is divided, the following set of characteristics will be calculated:
+\begin{equation}
+\label{eq:26}
+\overrightarrow{R}(i)=\{R_1(i), R_2(i), \dots, R_p(i)\},
+\end{equation}
+where
+\begin{equation}
+\label{eq:27}
+R_j(i)=m_j \varrho_i+\frac{(z_i(j)-z_{i-1}(j))^2}{m_j \varrho_i} - 2(z_{i-1}(j)+z_i(j)),1 \leq i < k,1 \leq j \leq k,
+\end{equation}
+\begin{equation}
+\label{eq:28}
+\varrho_i=(x_i-x_{i-1} )^{1/N}  ,1 \leq i<k,
+\end{equation}
+\begin{equation}
+\label{eq:29}
+m_j = r M_j, M_j=\frac{\max|z_i (j)-z_{i-1} (j)|}{\varrho_i} ,1\leq i < k, 1 \leq j \leq p.
+\end{equation}
+($m_j$, $1 \leq j \leq p$, is the estimate of the H{\"o}lder  constant in the condition (\ref{eq:10}) for the function $F(\lambda_j,y)$, $1 \leq j \leq p$, of the set $\Phi_t(y)$ from (\ref{eq:21})).
+
+The algorithm for simultaneous optimization of all functions of the set $\Phi_t(y)$ from (\ref{eq:21}) (further denoted as SGSA) can be represented as the following sequence of steps.
+
+\textit{Step 1.} Compute characteristics of the intervals $R_j (i)$, $1 \leq i < k$, $1 \leq i \leq p$, and determine the function $F(\lambda_j,y)$, $1 \leq j \leq p$, whose search information contains the interval with the maximum characteristic
+\begin{equation}
+\label{eq:30}
+R_q (t)=\max R_j (i), 1 \leq i < k, 1 \leq i \leq p.
+\end{equation}
+
+\textit{Step 2.} Select the point of the next iteration in the interval with the maximum characteristic
+\begin{equation}
+\label{eq:31}
+x^{k+1}=X(x_{t-1},x_t )
+\end{equation}
+and calculate the value $\overrightarrow{z}^{k+1}$ of all simultaneously optimized functions $F(\lambda_j,y)$, $1 \leq j \leq p$, at the point $x^{k+1}$ (when calculating the point  $x^{k+1}$ the values $z_i (q)$, $1 \leq i \leq k$, of the function $F(\lambda_q,y(x_i))$ whose number was determined at Step 1 should be used).
+
+\textit{Step 3.} Check the stopping condition according to (\ref{eq:18})
+\begin{equation}
+\label{eq:32}
+\varrho_t \leq \varepsilon
+\end{equation}
+
+When the functions $F(\lambda_j,y)$, $1 \leq j \leq p$, are optimized simultaneously, it should be kept in mind that the values of these functions at their global minima may differ.
+To ensure convergence to global minima of all the functions being optimized simultaneously, the SGSA algorithm has to be supplemented by a preliminary step of homogenizing the functions $F(\lambda_j,y)$, $1 \leq j \leq p$.
+
+\textit{Step 0.} Convert $F(\lambda_j,y)$, $1 \leq j \leq p$, according to the rule
+\begin{equation}
+\label{eq:33}
+F'(\lambda_j,y)=(F(\lambda_j,y)-z_{min} (j))/H_j, 1 \leq j \leq p,
+\end{equation}
+where $z_{min}(j)$, $1 \leq j \leq p$, is the minimum value of the function $F(\lambda_j,y)$, $1 \leq j \leq p$, i.e. 
+\begin{equation}
+\label{eq:34_2}
+z_{min} (j)=\min_{y\in D}F(\lambda_j,y), 1 \leq j \leq p,
+\end{equation}
+and $H_j$, $1 \leq j \leq p$, is the H{\"o}lder constant estimate in the condition (\ref{eq:10}) for the function $F(\lambda_j,y)$, $1 \leq j \leq p$.
+
+In the case where the values of $z_{min} (j)$, $H_j$, $1 \leq j \leq p$, are not known a prori, these values can be replaced by estimates calculated on the basis of available search information $A_k$ from (\ref{eq:22}) according to expressions (\ref{eq:19}) and (\ref{eq:29}).
+
+\section{Section} \label{sec:05}
+
 %
@@
 %
- \bibliographystyle{splncs04}
- \bibliography{bibliography}
+\begin{thebibliography}{1}
+\providecommand{\url}[1]{\texttt{#1}}
+\providecommand{\urlprefix}{URL }
+\providecommand{\doi}[1]{https://doi.org/#1}
+
+\bibitem{c1} Miettinen, K.: Nonlinear Multiobjective Optimization. Springer (1999).
+\bibitem{c2} Ehrgott, M.: Multicriteria Optimization. Springer (2nd ed., 2010).
+\bibitem{c3} Collette, Y., Siarry, P.:  Multiobjective Optimization: Principles and Case Studies (Decision Engineering). Springer (2011).
+\bibitem{c4} Marler, R. T., Arora, J. S.: Multi-Objective Optimization: Concepts and Methods for Engineering. VDM Verlag (2009).
+\bibitem{c5} Pardalos, P.M., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-Convex Multi-Objective Optimization. Springer (2017).
+\bibitem{c6} Marler, R. T., Arora, J. S. Survey of multi-objective optimization methods for engineering. Struct. Multidisciplinary Optimization \textbf{26}, 369--395 (2004).
+\bibitem{c7}  Figueira, J., Greco, S., Ehrgott, M., eds.: Multiple criteria decision analysis: State of the art surveys. New York (NY). Springer (2005).
+\bibitem{c8} Zavadskas, E. K., Turskis, Z., Kildien{\. e}, S.: State of art surveys of overviews on MCDM/MADM methods. Technological and Economic Development of Economy, \textbf{20}, 165--179 (2014).
+\bibitem{c9} Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications. Surv. Math. Ind. \textbf{11}, 1--42 (2005).
+\bibitem{c10} Collette, Y., Siarry, P.: Multiobjective Optimization: Principles and Case Studies. Decision Engineering. Springer, Heidelberg (2011). \doi{10.1007/978-3-662-08883-8}
+\bibitem{c11} Eichfelder, G.: Scalarizations for adaptively solving multi-objective optimization problems. Comput. Optim. Appl. \textbf{44}, 249--273 (2009).
+\bibitem{c12} Figueira, J., Liefooghe, A., Talbi, E., Wierzbicki, A.: A parallel multiple reference point approach for multi-objective optimization.
+European Journal of Operational Research, \textbf{205}(2), 390--400 2010.
+\doi{10.1016/j.ejor.2009.12.027}
+\bibitem{c13} Zhigljavsky, A.A.: Theory of Global Random Search. Dordrecht: Kluwer Academic Publishers (1991).
+\bibitem{c14} Pint{\' e}r, J.D.: Global optimization in Action (continuous and Lipschitz optimization: algorithms, implementations and applications). Kluwer Academic Publishers, Dortrecht (1996).
+\bibitem{c15} Strongin, R., Sergeyev, Ya.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2nd ed. 2013, 3rd ed. 2014).
+\bibitem{c16} Yang, X.-S.: Nature-inspired metaheuristic algorithms. Luniver Press, Frome (2008).
+\bibitem{c17} Locatelli, M., Schoen, F.: Global Optimization: Theory, Algorithms, and Applications. SIAM (2013).
+\bibitem{c18} Paulavi{\v c}ius, R., {\v Z}ilinskas, J.: Simplicial Global Optimization. Springer, New York (2014). 
+\bibitem{c19} Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. Princeton University Press (2016).
+\bibitem{c20} Sergeyev, Y.D., Kvasov, D.E.: Deterministic Global Optimization: An Introduction to the Diagonal Approach. Springer, New York (2017). 
+\bibitem{c21} Sergeyev Y.D., Strongin R.G., Lera D.: Introduction to global optimization exploiting space-filling curves. Springer (2013).
+\bibitem{c22} Piyavskij, S.: An algorithm for finding the absolute extremum of a function, Computational Mathematics and Mathematical Physics \textbf{12}, 57--67 (1972).
+\bibitem{c23} Shubert, B.O.: A sequential method seeking the global maximum of a function. SIAM Journal on Numerical Analysis \textbf{9} , 379--388 (1972).
+\bibitem{c24} Strongin R.G.: Multiextremal Minimization. Autom. Remote Control. \textbf{7}, 1085--1088 (1970).
+\bibitem{c25} Galperin, E.A.: The cubic algorithm. Journal of Mathematical Analysis and Applications \textbf{112}, 635--640 (1985).
+\bibitem{c26} Breiman, L., Cutler, A.: A deterministic algorithm for global optimization, Mathematical Programming \textbf{58}, 179--199 (1993).
+\bibitem{c27} Baritompa, W.: Accelerations for a variety of global optimization methods. Journal of Global Optimization \textbf{4}, 37--45 (1994).
+\bibitem{c28} Gergel, V.P.: A method of using derivatives in the minimization of multiextremum functions. Computational Mathematics and Mathematical Physics, \textbf{36}(6), 729--742 (1996).
+\bibitem{c29} Sergeyev, Y.D.: Global one-dimensional optimization using smooth auxiliary functions. Mathematical Programming \textbf{81}, 127--146 (1998).
+\bibitem{c30} Sergeyev Ya.D., Grishagin V.A.: Sequential and parallel global optimization algorithms. Optimization Methods and Software, \textbf{3}, 111--124 (1994).
+\bibitem{c31} Sergeyev, Y.D.: An information global optimization algorithm with local tuning. SIAM J. Optim., \textbf{5}(4), 858--870 (1995).
+\bibitem{c32} Sergeyev, Y.D., Grishagin, V.A.: Parallel asynchronous global search and the nested optimization scheme. J. Comput. Anal. Appl., \textbf{3}(2), 123--145 (2001).
+\bibitem{c33} Lera D., Sergeyev Ya.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and Holder constants.
+Communications in Nonlinear Science and Numerical Simulation, \textbf{23}, 328--342 (2015).
+\bibitem{c34} Grishagin, V., Israfilov, R., Sergeyev, Y.: Comparative efficiency of dimensionality reduction schemes in global optimization. AIP Conference Proceedings, 1776 (2016).
+\bibitem{c35} Gergel, V.: An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems.
+2nd International Conference on Mathematics and Computers in Sciences and in Industry, 13--18 (2015).
+\doi{10.1109/MCSI.2015.18}
+\bibitem{c36} Gergel, V.P., Kozinov, E.A.: Accelerating multicriterial optimization by the intensive exploitation of accumulated search data.
+AIP Conference Proceedings, \textbf{1776}, 090003 (2016).
+\doi{10.1063/1.4965367}
+\bibitem{c37} Gergel, V., Kozinov, E. Efficient multicriterial optimization based on intensive reuse of search information. Journal of Global Optimization, 2018, 71(1), 73--90.
+\bibitem{c38} Gergel, V., Barkalov, K., Sysoyev, A.
+Globalizer: A novel supercomputer software system for solving time-consuming global optimization problem.
+Numerical algebra, control and optimization, \textbf{8}(1), 47--62 (2018).
+\bibitem{c39} {\v Z}ilinskas, A., Zilinskas, J.: Adaptation of a one-step worst-case optimal univariate algorithm of bi-objective Lipschitz optimization to multidimensional problems.
+Commun Nonlinear Sci Numer Simulat, \textbf{21}(1-3), 89--98 (2015).
+\doi{10.1016/j.cnsns.2014.08.025}
+\bibitem{c40} Evtushenko.
+Y., Posypkin, M.: A deterministic algorithm for global multiobjective optimization, Optimization Methods \& Software, \textbf{29}(5), 1005--1019 (2014).
+\doi{10.1080/10556788.2013.854357}
+\bibitem{c41} Evtushenko, Y., Posypkin, M.: Method of non-uniform coverages to solve the multicriteria optimization problems with guaranteed accuracy.
+Autom.
+Remote Control, \textbf{75}(6), 1025--1040 (2014).
+\bibitem{c42} Bleuler, S., Laumanns, M., Thiele, L., Zitzler, E.: Pisa-a platform and programming language independent interface for search algorithms, Evolutionary Multi-Criterion Optimization.
+LNCS, \textbf{2632}, 494--508 (2003).
+\doi{10.1007/3-540-36970-8\_35}
+\bibitem{c43} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
+ACM Transactions on Mathematical Software, \textbf{29}(4), 469--480 (2003).
+\end{thebibliography}
+
 %

### 2021-02-12 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
1. Добавлен раздел с экспериментами и заключение

@@
 \usepackage{graphicx}
+\usepackage{multirow}
 % Used for displaying a sample figure. If possible, figure files should
@@
 
-\section{Section} \label{sec:05}
+\section{Results of numerical experiments} \label{sec:05}
+
+Numerical experiments were performed on the Lobachevsky supercomputer (University of Nizhny Novgorod), (operating system -- CentOS 6.4, control system -- SLURM).
+One supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM.
+The CPU is 8-core (i.e.
+a total of 16 CPU cores are available on the node).
+The numerical experiments were performed using the Globalizer system \cite{c36}.
+
+      The first series of experiments was performed to compare the MGSA algorithm with a number of well-known multicriteria optimization algorithms by solving a bi-criteria test problem \cite{c40}:
+\begin{equation}
+\label{eq:35}
+f_1 (y)=(y_1-1) y_2^2+1,f_2 (y)=y_2, 0 \leq y_1, y_2 \leq 1.
+\end{equation}
+
+In the course of the experiments, a numerical approximation of the Pareto domain was carried out to solve the problem (\ref{eq:35}), and the quality of the approximation was evaluated using the hypervolume and distribution uniformity indices (HV) \cite{c37,c40}.
+The first of these indices characterizes the completeness of approximation (a larger value corresponds to a more complete coverage of the Pareto domain), while the second one shows the uniformity of coverage (a smaller value corresponds to a more uniform coverage of the Pareto domain).
+
+Five multicriteria optimization algorithms were compared in this experiment: the Monte-Carlo (MC) method, the genetic algorithm SEMO from the PISA library \cite{c42}, the non-uniform coverage (NUC) method \cite{c40}, the bi-objective Lipschitz optimization (BLO) method \cite{c39} and the MGSA algorithm proposed in this paper.
+For the first three algorithms, the numerical results were used from \cite{c41}.
+The results of the BLO method were presented in \cite{c39}.
+
+For MGSA 100 problems (\ref{eq:04}) with different values of convolution coefficients $\lambda$ uniformly distributed in $\Lambda$ were solved.
+The full results of the experiments performed are presented in Table \ref{tab:1}.
+
+\begin{table}[ht]
+\centering
+\caption{Comparison of the efficiency of multicriteria optimization algorithms}
+\label{tab:1}
+\begin{tabular}{cccccc}\hline
+\textbf{Solution method}                                                                  & \textbf{MC} & \textbf{SEMO} & \textbf{NUC} & \textbf{BLO} & \textbf{MGSA}  \\ \hline
+\begin{tabular}[c]{@{}c@{}}Number of \\ method iterations\end{tabular}                    & 500         & 500           & 515          & 498          & \textbf{338}   \\
+\begin{tabular}[c]{@{}c@{}}Number of \\ points found in \\ the Pareto domain\end{tabular} & 67          & 104           & 29           & 68           & \textbf{115}  \\
+HV index                                                                                  & 0.300       & 0.312         & 0.306        & 0.308        & \textbf{0.318} \\
+DU index                                                                                  & 1.277       & 1.116         & 0.210        & 0.175        & \textbf{0.107} \\ \hline
+\end{tabular}
+\end{table}
+
+As the experimental results show, the MGSA algorithm has a distinct advantage over the other multicriteria optimization methods considered, even for solving relatively simple MOO problems.
+
+   In the second series of numerical experiments, we solved bi-criteria two-dimensional MOO problems, i.e.
+with $N = 2$, $s = 2$.
+We used multiextremal functions obtained with the help of the GKLS generator \cite{c43} as objective functions of the problem.
+During the experiments, we solved 100 multicriteria problems of this class, for each of them the set $\Phi_p (y)$ from (\ref{eq:21}) comprised 5, 10 and 25 simultaneously optimized functions respectively (the convolution coefficients $\lambda \in \Lambda$ from (\ref{eq:06}) for the functions of the set $\Phi_p (y)$ being uniformly distributed in $\Lambda$).
+To check the accuracy of the solution of MOO problems, the computed estimates of the efficient solutions were checked to confirm they belonged to the Pareto domain.
+For the parameters of the MGSA algorithm, the following values were used: accuracy $\varepsilon=0.01$,  reliability $r=5.6$.
+The computational results were averaged over the number of MOO problems solved.
+
+The results of numerical experiments are presented in Table \ref{tab:2}.
+
+\begin{table}[ht]
+\centering
+\caption{Results of numerical experiments on solving two-dimensional bi-criteria MOO problems}
+\label{tab:2}
+\begin{tabular}{cccclccclccc}
+\hline
+\multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}Number of \\ functions in\\ $\Phi_p (y)$\end{tabular}} & \multicolumn{11}{c}{Algorithms}                                                     \\ \cline{2-12} 
+                                                                                        & \multicolumn{3}{c}{GSA} &  & \multicolumn{3}{c}{MGSA} &  & \multicolumn{3}{c}{SGSA} \\
+                                                                                        & Iters   & HV    & DU    &  & Iters   & HV     & DU    &  & Iters   & HV     & DU    \\ \hline
+5                                                                                       & 3771,8  & 6,455 & 0,25  &  & 1823,4  & 6,481  & 0,208 &  & 2090,5  & 6,479  & 0,217 \\
+10                                                                                      & 7941,6  & 6,485 & 0,183 &  & 1955,5  & 6,486  & 0,205 &  & 2224,4  & 6,486  & 0,191 \\
+25                                                                                      & 20456,2 & 6,504 & 0,143 &  & 2135,8  & 6,490  & 0,209 &  & 2456,1  & 6,493  & 0,188 \\ \hline
+\end{tabular}
+\end{table}
+
+Table 2 shows in the first column the number of functions $\Phi_p (y)$ being optimized from (\ref{eq:21}).
+The columns named Iters contain the average number of iterations performed by the MOO algorithm to solve the MOO problem.
+The columns named HV and DU contain the values of the HV and DU indicators.
+The results of the numerical experiments are divided with respect to the three algorithms used: GSA is described in Section \ref{sec:03}, MGSA is discussed in Section \ref{sec:03} (item 2), SGSA is presented in Section \ref{sec:04} (item 3).
+
+The experimental results show that by reusing search information when optimizing 25 functions from the set $\Phi_p (y)$ the total computational volume can be reduced  by more than 9.5 times without resorting to any additional computational resources.
+At the same time, the MGSA algorithm performs the least number of iterations (the number of calculations of objective function values), while the SGSA algorithm provides a better approximation of the Pareto domain both in terms of approximation completeness and coverage uniformity.
+It is also noteworthy that the average number of iterations to optimize one function from the set $\Phi_p (y)$ decreases by more than 4 times as the number of functions in the set $\Phi_p (y)$ increases -- see Table \ref{tab:3}.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+\begin{table}[ht]
+\centering
+\caption{Average number of iterations required to optimize one function from the set $\Phi_p (y)$}
+\label{tab:3}
+\begin{tabular}{cccc}
+\hline
+\multirow{2}{*}{\begin{tabular}[c]{@{}c@{}}Number of \\ functions in $\Phi_p (y)$\end{tabular}} & \multicolumn{3}{c}{Average number of   iterations} \\
+                                                                                        & GSA             & MGSA            & SGSA           \\ \hline
+5                                                                                       & 754,4           & 364,7           & 418,1          \\
+10                                                                                      & 794,2           & 195,6           & 222,4          \\
+25                                                                                      & 818,2           & 85,4            & 98,2           \\ \hline
+\end{tabular}
+\end{table}
+
+In the third series of numerical experiments, 10 bi-criteria five-dimensional MOO problems were solved, i.e.
+$N = 5$, $s = 2$, $m=2$.
+The objective functions of the MOO problems were determined, as before, using the GKLS generator \cite{c43}.
+When solving the problems, the accuracy $\varepsilon=0.05$ and the reliability of the method $r=5.6$ were used.
+The results of numerical experiments are presented in Table \ref{tab:4}.
+
+\begin{table}[ht]
+\centering
+\caption{The results of a series of experiments to solve five-dimensional bi-criteria MOO problems}
+\label{tab:4}
+\begin{tabular}{cccclccclccc}
+\hline
+\multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}Number of \\ functions in\\ $\Phi_p (y)$\end{tabular}} & \multicolumn{11}{c}{Algorithms}                                                     \\ \cline{2-12} 
+                                                                                        & \multicolumn{3}{c}{GSA} &  & \multicolumn{3}{c}{MGSA} &  & \multicolumn{3}{c}{SGSA} \\
+                                                                                        & Iters   & HV    & DU    &  & Iters   & HV     & DU    &  & Iters   & HV     & DU    \\ \hline
+25                                                                                      & 1030039,6 & 2,440 & 0,543 & & 619755,6 & 2,441 & 0,681 & & 669743,8 & 2,364 & 0,668 \\ \hline
+\end{tabular}
+\end{table}
+
+The results of our experiments show that with increasing dimensionality of the MOO problems to be solved, the trend that has been identified continues: the amount of computation (the number of global search iterations) is reduced by more than 1.6 times due to reuse of search information, the MGSA algorithm performs the least number of iterations, and the SGSA algorithm provides the best approximation of the Pareto domain.
+
+\section*{Conclusion}
+
+This paper proposes an effective method for solving computationally intensive multicriteria optimization problems which require computing several Pareto-optimal decisions.
+It is also assumed that efficiency criteria may be multiextremal, and computation of function values may require a large amount of calculations.
+The proposed approach is based on the reduction of multicriteria optimization problems to one-dimensional global optimization problems that can be solved using effective statistical information algorithms of global search.
+The novelty of the proposed approach consists in the possibility of solving several global optimization problems simultaneously, which allows multiple Pareto-optimal decisions to be obtained.
+Besides, such approach provides for reuse of the computed search information, which considerably reduces computational effort for solving multicriteria optimization problems.
+
+The results of numerical experiments show that by using the developed approach it is possible to achieve a significant reduction of computational complexity when solving multicriteria optimization problems.
+
+It should be noted in conclusion that the proposed approach is quite promising and requires further research in this area.
+First of all, it is necessary to continue numerical experiments on solving multicriteria optimization problems with a greater number of partial efficiency criteria and for higher dimensionality of optimization problems to be solved.
+The possibility of organizing parallel computations for high-performance systems should also be evaluated.
 

### 2021-02-13 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
Motor исправление текста статьи

@@
 \begin{abstract}
-This paper considers computationally intensive multicriteria optimization problems which require computing multiple Pareto-optimal decisions.
+This paper considers computationally intensive multi-objective optimization problems which require computing multiple Pareto-optimal decisions.
 It is also assumed that efficiency criteria may be multiextremal, and the cost of calculating function values may be quite high.
-The proposed approach is based on the reduction of multicriteria optimization problems to one-dimensional global optimization problems that can be solved using effective statistical information algorithms of global search.
+The proposed approach is based on the reduction of multi-objective optimization problems to one-dimensional global optimization problems that can be solved using effective information-statistical algorithms of global search.
 One key innovation of the developed approach consists in the possibility of solving several global optimization problems simultaneously, which allows multiple Pareto-optimal decisions to be obtained.
-Besides, such approach provides for reuse of the computed search information, which considerably reduces computational effort for solving multicriteria optimization problems.
+Besides, such approach provides for reuse of the computed search information, which considerably reduces computational effort for solving multi-objective optimization problems.
 Computational experiments confirm the potential of the proposed approach.
@@
 
-Choosing the best options (decision making) in situations with many different alternatives is a problem that occurs in almost every domain of human activity.
+Choosing the optimal options (decision making) in situations with many different alternatives is a problem that occurs in almost every domain of human activity.
 In the simplest cases, decision making problems can be viewed as optimization problems of various kinds such as convex programming, discrete optimization, nonlinear programming, etc.
@@
 In the most general case, however, the effectiveness of decisions may be determined by several objective criteria, which necessitates solving multi-objective optimization (MOO) problems.
-Thus, MOO problems are the most common decision making settings that one has to solve in many scientific and technical applications.
+Thus, MOO problems are the most common decision making statements that one has to solve in many scientific and technical applications.
 A large number of approaches and methods have been developed to solve such problems and they have been used to solve many decision making problems in various fields of practical applications -- see e.g.
@@
 
-One of the approaches most commonly used in the search for efficient solutions is scalarization of the vector efficiency criterion, when a MOO problem is reduced to solving one or several scalar (in the general case, global) optimization problems in which the objective function is the only function.
+One of the approaches most commonly used in the search for efficient decisions is scalarization of the vector efficiency criterion, when a MOO problem is reduced to solving one or several scalar (in the general case, global) optimization problems in which the objective functions are converted to the integrated scalar function.
 The family of optimization problems generated in such approach is further referred to as multiple global optimization problem (MGO).
@@
 The structure of the paper is as follows.
-In Section \ref{sec:02}, we give the statement of the multicriteria optimization problem.
-Section \ref{sec:03} presents a scheme for reduction of multicriteria optimization problems to one-dimensional global optimization problems as well as some methods for solving such problems.
+In Section \ref{sec:02}, we give the statement of the multi-objective optimization problem.
+Section \ref{sec:03} presents a scheme for reduction of multi-objective optimization problems to one-dimensional global optimization problems as well as some methods for solving such problems.
 Section \ref{sec:04} considers the proposed approach for simultaneous solution of several global optimization problems yielding several Pareto-optimal decisions at a time.
@@
 It is also assumed that $f_i(y)$, $1 \leq i \leq s$, are multiextremal and have the form of time-consuming ``black-box'' computational procedures.
-It is also assumed that the objective functions.
 It is also assumed that the objective functions $f_i(y)$, $1 \leq i \leq s$, satisfy the Lipschitz condition
@@
 where $F$ is a scalar multiextremal function generated as a result of scalarization of objective functions $f_i$, $1 \leq i \leq s$, $\lambda$ is the vector of parameters of the applied convolution of functions, and $D$ is the search domain from (\ref{eq:01}).
-In the proposed approach, we use for scalarization the compromise programming method \cite{c2,c11}, where solution of the MOO problem consists in finding the efficient decision corresponding most closely to the optimality measures of the specified reference decision $y^0 \in D$.
-In this case, a possible variant of convolution of objective functions $f_i$, $1 \leq i \leq s$, can have the form
+In the proposed approach, we use for scalarization the compromise programming method \cite{c2,c11}, where solution of the MOO problem consists in finding the efficient decision corresponding most closely to the values of the objective functions of the specified reference decision $y^0 \in D$.
+In this case, a possible statement of the statement scalar optimization problem can have the form
 \begin{equation}
 \label{eq:05}
-\min F(\lambda,y)=1/s \sum_{i=1}^s {\lambda_i (f_i (y)-f_i (y^0 ))^2}, y \in D,
+\min F(\lambda,y)=\min 1/s \sum_{i=1}^s {\lambda_i (f_i (y)-f_i (y^0 ))^2}, y \in D,
 \end{equation}
-where $F(\lambda,y)$ is the standard deviation of the values of the objective functions $f_i$, $1 \leq i \leq s$, for the decision $y \in D$ and for the specified reference decision $y^0 \in D$, while the coefficients $0 \leq \lambda_i \leq 1$, $1 \leq i < s$, are measures of the significance of the approximation accuracy for each variable $y_i$, $1 \leq i \leq N$, separately  without loss of generality, we can assume that the area of possible values of the coefficients $\lambda$ is a set
+where $F(\lambda,y)$ is the standard deviation of the values of the objective functions $f_i$, $1 \leq i \leq s$, for the decision $y \in D$ and for the specified reference decision $y^0 \in D$, while the coefficients $0 \leq \lambda_i \leq 1$, $1 \leq i < s$, are measures of the significance of the approximation accuracy for each variable $y_i$, $1 \leq i \leq N$, separately.
+Without loss of generality, we can assume that the area of possible values of the coefficients $\lambda$ is a set
 \begin{equation}
@@
 The reference decision $y^0 \in D$ in (\ref{eq:05}) can be known \textit{a priori} or determined on the basis of some known prototype.
-In many cases, an abstract ideal decision $y^0 \in D$ is used as the reference decision $y^*\in D$ in which the objective functions $f_i$, $1 \leq i \leq s$, assume minimal possible values, i.e.
+In many cases, an abstract ideal decision $y^0 \in D$ is used as the reference decision $y^*\in D$ in which the objective functions $f_i$, $1 \leq i \leq s$, have minimal possible values, i.e.
 \begin{equation}
@@
 \end{equation}
-
 where the constant $H$ is defined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:08}), and $N$ is the dimensionality of the optimization problem (\ref{eq:01}).
-
-As a result of dimensionality reduction, the computational scheme for solving problem (\ref{eq:09}) consists of the following (see Fig. \ref{fig:1}):
-
-\begin{itemize}
-	\item The optimization algorithm minimizes the reduced one-dimensional function $\varphi(y(x))$ from (\ref{eq:09}),
-	\item After determining the point $x \in [0,1]$ of the next iteration of the one-dimensional global search, the multi-dimensional image $y \in D$ is computed for the mapping $y(x)$,
-	\item At the multi-dimensional point $y \in D$, the value of the original multi-dimensional function $\varphi(y)$ is computed from (\ref{eq:04}),
-	\item The computed value $z = \varphi(y)$ is then used as the value of the reduced one-dimensional function $\varphi(y(x))$ at  the point $x\in[0,1]$.
-\end{itemize}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{fig1}
-  \caption{General scheme for solving the problem (\ref{eq:09})}
-  \label{fig:1}
-\end{figure}
 
@@
 \cite{c22,c23,c24,c25,c26,c27,c28,c29}.
-At the same time, it should be noted that most works where dimensionality reduction is a key feature for solving multiextremal problems like (\ref{eq:04}) rely on the statistical information theory of global search \cite{c15}.
-This theory has provided the a basis for developing a large number of effective methods for multi-extremal optimization -- see, e.g., \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
+At the same time, it should be noted that most works where dimensionality reduction is a key feature for solving multiextremal problems like (\ref{eq:04}) rely on the information-statistical theory of global search \cite{c15}.
+This theory has provided the a basis for developing a large number of efficient methods for multi-extremal optimization -- see, e.g., \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
 
-Within the framework of statistical information theory, a general computational scheme of global optimization algorithms was proposed which in brief is as follows \cite{c15,c30,c35}.
+Within the framework of information-statistical theory, a general computational scheme of global optimization algorithms was proposed which in brief is as follows \cite{c15,c30,c35}.
 
@@
 This estimate is determined by means of characteristics of intervals $R(i)$, $1 \leq i<k$, whose values should be proportional to the degree of possibility that the global minimum is located  in these intervals.
-The type of these characteristics depends on the global optimization algorithm used -- thus, for example, for an algorithm which constructs a uniform dense grid in the global search area, the characteristic is simply the interval length
+The type of these characteristics depends on the global optimization algorithm used -- thus, for example, for an algorithm which constructs a uniform dense grid in the global search domain, the characteristic is simply the interval length
 \begin{equation}
@@
 where $H$ is the H{\" o}lder constant from (\ref{eq:10}) for the reduced global optimization problem being solved (\ref{eq:09}), $z_i = \varphi(y(z_i))$, $1 \leq i < k$ and $N$ is the dimensionality of the problem from (\ref{eq:01})).
-For the global search algorithm (GSA) \cite{c15,c24} developed in the framework of the statistical information approach, the characteristic is
+For the global search algorithm (GSA) \cite{c15,c24} developed in the framework of the information-statistical approach, the characteristic is
 \begin{equation}
@@
 \label{eq:17}
-x^{k+1}=X(x_{t-1}, x_t),
+x^{k+1}=X(x_{t-1}, x_t)\footnote{The rule $X$ for selecting the point $x^{k+1}$ of the next iteration in the interval $(x_{t-1}, x_t)$ is stated by the global optimization algorithm},
 \end{equation}
@@
 where $\varrho_t = (x_t-x_{t-1})^{1⁄N}$, $t$ from (\ref{eq:16}) and $\varepsilon > 0$ is the specified accuracy of the solution.
-If the stopping condition (\ref{eq:18}) is reached, then the solving of the optimization problem is stopped, otherwise $k=k+1$ is assumed and the next global search iteration begins.
+If the stopping condition (\ref{eq:18}) is met, then the solving of the optimization problem is stopped, otherwise $k=k+1$ is assumed and the next global search iteration begins.
 
@@
 It should be noted again that the computational scheme discussed above is quite general.
-Many global search algorithms can be represented within this scheme, as evidenced, in particular, by examples (\ref{eq:12})-(\ref{eq:14}) and other characteristic algorithms -- see, for example, \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
+Many global search algorithms can be represented within this scheme, as evidenced, in particular, by examples (\ref{eq:12})-(\ref{eq:14}) and other global optimization algorithms -- see, for example, \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
 
-Convergence conditions for characteristic algorithms depend on the properties of interval characteristics used.
-One of sufficient conditions for convergence of algorithms is, for example, the requirement that the characteristic of the interval containing the global minimum point should take on the maximum value at step 1 of the characteristic scheme during the global search iterations.
+Convergence conditions of the algorithms formulated in the form (\ref{eq:16})-(\ref{eq:18}).
+One of sufficient conditions for convergence of algorithms is, for example, the requirement that the characteristic of the interval containing the global minimum point should take on the maximum value at step 1 of the scheme (\ref{eq:16})-(\ref{eq:18}) during the global search iterations.
 This condition is satisfied, for example, for the multidimensional generalized algorithms proposed in \cite{c22,c23} when the H{\"o}lder constant from (\ref{eq:10}) is specified exactly.
@@
 
-\section{An Approach for Simultaneous Finding of Multiple Effective Solutions in Multi-objective Optimization Problems} \label{sec:04}
+\section{An Approach for Simultaneous Finding of Multiple Efficient Decisions in Multi-objective Optimization Problems} \label{sec:04}
 
 \textbf{1.
-The need to solve multiple global optimization problems.} As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different effective decisions due to possible changes in the optimality requirements.
-Obtaining different  effective decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multi-extremal functions $F(\lambda, y)$ from (\ref{eq:04}).
-The resulting different functions $F(\lambda, y)$ for different values of the coefficients $\lambda \in \Lambda$ can be optimized sequentially.
-This deermines a multi-stage scheme for solving the MOO problem, when at each stage DM specifies the necessary scalarization coefficients, then the resulting scalar optimization problem is solved, after which DM analyzes the effective solution found.
+The need to solve multiple global optimization problems.} As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different efficient decisions due to possible changes in the optimality requirements.
+Obtaining different efficient decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multi-extremal functions $F(\lambda, y)$ from (\ref{eq:04}).
+These functions $F(\lambda, y)$ for different values of the coefficients $\lambda \in \Lambda$ can be optimized sequentially.
+This determines a multi-stage scheme for solving the MOO problem, when at each stage DM specifies the necessary scalarization coefficients, then the resulting scalar optimization problem is solved, after which DM analyzes the efficient decisions found.
 
@@
 Having such an option makes it easier for DM  to specify importance coefficients $\lambda \in \Lambda$ of the objective functions $f_i(y)$, $1 \leq i \leq s$.
-By solving simultaneously several generated scalar optimization problems, one can  obtain the estimates of efficient decisions at the earliest stages of computing, which makes it possible to dynamically (in the process of computation) change the set of problems being solved -- to stop solving those that obviously have no prospect of success (from DM's point of view) or to add new optimization problems.
+By solving simultaneously several generated scalar optimization problems, one can  obtain the estimates of efficient decisions at the earliest stages of computing, which makes it possible to dynamically (in the process of computations) change the set of problems being solved -- to stop solving those that obviously have no prospect of success (from DM's point of view) or to add new optimization problems.
 
@@
 \label{eq:21}
-\Phi_p (y)=\{ F(\lambda_1,y), F(\lambda_2,y),\dots,F(\lambda_p,y)  \},  \lambda_i \in \Delta, 1 \leq i \leq p
+\Phi_p (y)=\{ F(\lambda_1,y), F(\lambda_2,y),\dots,F(\lambda_p,y)  \},  \lambda_i \in \Lambda, 1 \leq i \leq p
 \end{equation}
@@
 \textbf{2.
-Step-by-step solution of a set of global optimization problems.} As shown in Section \ref{sec:03}, the statistical information multiextremal optimization algorithms used in the proposed approach determine the points of consecutive iterations of the global search taking into account the search information
+Step-by-step solution of a set of global optimization problems.} As shown in Section \ref{sec:03}, the information-statistical multiextremal optimization algorithms used in the proposed approach determine the points of consecutive iterations of the global search taking into account the search information
 \begin{equation}
@@
 
-It is important to note that since the set $\Phi_t(y)$ of functions being optimized  simultaneously is generated from the same MOO problem from (\ref{eq:01}), the existence of the set $A_k$ from (\ref{eq:12}) allows us to adjust the results of all previously performed calculations of the values of the objective functions $f_i(y)$, $1 \leq i \leq s$, to the values of the next optimized function $F(\lambda, y)$ from (\ref{eq:04}) without repeating any time consuming calculations of values, i.e.
+It is important to note that since the set $\Phi_p(y)$ of functions being optimized  simultaneously is generated from the same MOO problem from (\ref{eq:01}), the existence of the set $A_k$ from (\ref{eq:12}) allows us to adjust the results of all previously performed calculations of the values of the objective functions $f_i(y)$, $1 \leq i \leq s$, to the values of the next optimized function $F(\lambda, y)$ from (\ref{eq:04}) without repeating any time consuming calculations of values, i.e.
 \begin{equation}
@@
 
-Thus, all the search information $A_k$ from (\ref{eq:22}), recalculated according to (\ref{eq:23}), can be reused to continue solving the problems of the set $\Phi_t(y)$.
-Such a possibility provides a significant reduction in computations up to performing only a limited set of global search iterations (see the results of computational experiments in Section \ref{sec:03}).
+Thus, all the search information $A_k$ from (\ref{eq:22}), recalculated according to (\ref{eq:23}), can be reused to continue solving the problems of the set $\Phi_p(y)$.
+Such a possibility provides a significant reduction in computations up to performing only a limited set of global search iterations (see the results of numerical experiments in Section \ref{sec:03}).
 
-     This type of information connectivity of functions in the set $\Phi_t(y)$ from (\ref{eq:21}) makes it possible to generalize the computational scheme (\ref{eq:16})-(\ref{eq:18}) for solving a single global optimization problem for the case of optimizing the functions of the set $\Phi_t(y)$ from (\ref{eq:21}) by adding a preliminary step of the search information  transformation.
+     This type of information connectivity of functions in the set $\Phi_p(y)$ from (\ref{eq:21}) makes it possible to generalize the computational scheme (\ref{eq:16})-(\ref{eq:18}) for solving a single global optimization problem for the case of optimizing the functions of the set $\Phi_p(y)$ from (\ref{eq:21}) by adding a preliminary step of the search information  transformation.
 
-\textit{Step 0.} Adjust the state of search information $A_k$ from (\ref{eq:22}) to the values of the function $F(\lambda, y)$ from the set $\Phi_t(y)$ according to rule (\ref{eq:23}).
+\textit{Step 0.} Adjust the state of search information $A_k$ from (\ref{eq:22}) to the values of the function $F(\lambda, y)$ from the set $\Phi_p(y)$ according to rule (\ref{eq:23}).
 
-The GSA algorithm applied to optimize the functions of the set $\Phi_t(y)$ from (\ref{eq:21}) and using search information $A_k$ will be further referred to as the Multiple Global Search Algorithm (MGSA).
+The GSA algorithm applied to optimize the functions of the set $\Phi_p(y)$ from (\ref{eq:21}) and using the search information $A_k$ will be further referred to as the Multiple Global Search Algorithm (MGSA).
 
 \textbf{3.
-Simultaneous solution of a set of global optimization problems.} Information connectivity makes it possible to propose a more general scheme for simultaneous optimization of all the functions of the set $\Phi_t(y)$ from (\ref{eq:21}).
-In this case, search information $A_k$ from (\ref{eq:22}) will contain the computed values of all simultaneously optimized functions $F(\lambda_i, y)$, $1 \leq i \leq p$, i.e.
+Simultaneous solution of a set of global optimization problems.} Information connectivity makes it possible to propose a more general scheme for simultaneous optimization of all the functions of the set $\Phi_p(y)$ from (\ref{eq:21}).
+In this case, the search information $A_k$ from (\ref{eq:22}) will contain the computed values of all simultaneously optimized functions $F(\lambda_i, y)$, $1 \leq i \leq p$, i.e.
 \begin{equation}
@@
 \label{eq:27}
-R_j(i)=m_j \varrho_i+\frac{(z_i(j)-z_{i-1}(j))^2}{m_j \varrho_i} - 2(z_{i-1}(j)+z_i(j)),1 \leq i < k,1 \leq j \leq k,
+R_j(i)=m_j \varrho_i+\frac{(z_i(j)-z_{i-1}(j))^2}{m_j \varrho_i} - 2(z_{i-1}(j)+z_i(j)),1 \leq i < k,1 \leq j \leq p,
 \end{equation}
@@
 \end{equation}
-($m_j$, $1 \leq j \leq p$, is the estimate of the H{\"o}lder  constant in the condition (\ref{eq:10}) for the function $F(\lambda_j,y)$, $1 \leq j \leq p$, of the set $\Phi_t(y)$ from (\ref{eq:21})).
+($m_j$, $1 \leq j \leq p$, is the estimate of the H{\"o}lder  constant in the condition (\ref{eq:10}) for the function $F(\lambda_j,y)$, $1 \leq j \leq p$, of the set $\Phi_p(y)$ from (\ref{eq:21})).
 
-The algorithm for simultaneous optimization of all functions of the set $\Phi_t(y)$ from (\ref{eq:21}) (further denoted as SGSA) can be represented as the following sequence of steps.
+The algorithm for simultaneous optimization of all functions of the set $\Phi_p(y)$ from (\ref{eq:21}) (further denoted as SGSA) can be represented as the following sequence of steps.
 
@@
 \end{equation}
-and $H_j$, $1 \leq j \leq p$, is the H{\"o}lder constant estimate in the condition (\ref{eq:10}) for the function $F(\lambda_j,y)$, $1 \leq j \leq p$.
+and $H_j$, $1 \leq j \leq p$, is the H{\"o}lder constant in the condition (\ref{eq:10}) for the function $F(\lambda_j,y)$, $1 \leq j \leq p$.
 
@@
 
-\section{Results of numerical experiments} \label{sec:05}
+\section{Results of Numerical Experiments} \label{sec:05}
 
@@
 
-      The first series of experiments was performed to compare the MGSA algorithm with a number of well-known multicriteria optimization algorithms by solving a bi-criteria test problem \cite{c40}:
+      The first series of experiments was performed to compare the MGSA algorithm with a number of well-known multi-objective optimization algorithms by solving a bi-criteria test problem \cite{c40}
 \begin{equation}
@@
 
-In the course of the experiments, a numerical approximation of the Pareto domain was carried out to solve the problem (\ref{eq:35}), and the quality of the approximation was evaluated using the hypervolume and distribution uniformity indices (HV) \cite{c37,c40}.
+In the course of the experiments, a numerical approximation of the Pareto domain was carried out to solve the problem (\ref{eq:35}), and the quality of the approximation was evaluated using the hypervolume (HV) and distribution uniformity (DU) indices \cite{c37,c40}.
 The first of these indices characterizes the completeness of approximation (a larger value corresponds to a more complete coverage of the Pareto domain), while the second one shows the uniformity of coverage (a smaller value corresponds to a more uniform coverage of the Pareto domain).
 
-Five multicriteria optimization algorithms were compared in this experiment: the Monte-Carlo (MC) method, the genetic algorithm SEMO from the PISA library \cite{c42}, the non-uniform coverage (NUC) method \cite{c40}, the bi-objective Lipschitz optimization (BLO) method \cite{c39} and the MGSA algorithm proposed in this paper.
+Five multi-objective optimization algorithms were compared in this experiment: the Monte-Carlo (MC) method, the genetic algorithm SEMO from the PISA library \cite{c42}, the Non-Uniform Coverage (NUC) method \cite{c40}, the Bi-objective Lipschitz Optimization (BLO) method \cite{c39} and the MGSA algorithm proposed in this paper.
 For the first three algorithms, the numerical results were used from \cite{c41}.
@@
 \centering
-\caption{Comparison of the efficiency of multicriteria optimization algorithms}
+\caption{Comparison of the efficiency of multi-objective optimization algorithms}
 \label{tab:1}
@@
 
-As the experimental results show, the MGSA algorithm has a distinct advantage over the other multicriteria optimization methods considered, even for solving relatively simple MOO problems.
+As the experimental results show, the MGSA algorithm has a distinct advantage over the other multi-objective optimization methods considered, even for solving relatively simple MOO problems.
 
@@
 with $N = 2$, $s = 2$.
-We used multiextremal functions obtained with the help of the GKLS generator \cite{c43} as objective functions of the problem.
-During the experiments, we solved 100 multicriteria problems of this class, for each of them the set $\Phi_p (y)$ from (\ref{eq:21}) comprised 5, 10 and 25 simultaneously optimized functions respectively (the convolution coefficients $\lambda \in \Lambda$ from (\ref{eq:06}) for the functions of the set $\Phi_p (y)$ being uniformly distributed in $\Lambda$).
+We used multiextremal functions obtained with the help of the GKLS generator \cite{c43} as objective functions of the MOO problem.
+During the experiments, we solved 100 multi-objective problems of this class, for each of them the set $\Phi_p (y)$ from (\ref{eq:21}) comprised 5, 10 and 25 simultaneously optimized functions respectively (the convolution coefficients $\lambda \in \Lambda$ from (\ref{eq:06}) for the functions of the set $\Phi_p (y)$ being uniformly distributed in $\Lambda$).
 To check the accuracy of the solution of MOO problems, the computed estimates of the efficient solutions were checked to confirm they belonged to the Pareto domain.
@@
 
+Table 2 shows in the first column the number of functions $\Phi_p (y)$ being optimized from (\ref{eq:21}).
+The columns named Iters contain the average number of iterations performed by the algorithm to solve the MOO problem.
+The columns named HV and DU contain the values of the HV and DU indicators.
+The results of the numerical experiments are divided with respect to the three algorithms used: GSA is described in Section \ref{sec:03}, MGSA is discussed in Section \ref{sec:03} (item 2), SGSA is presented in Section \ref{sec:04} (item 3).
+
 \begin{table}[ht]
 \centering
-\caption{Results of numerical experiments on solving two-dimensional bi-criteria MOO problems}
+\caption{Results of numerical experiments on solving bi-criteria two-dimensional MOO problems}
 \label{tab:2}
@@
 
-Table 2 shows in the first column the number of functions $\Phi_p (y)$ being optimized from (\ref{eq:21}).
-The columns named Iters contain the average number of iterations performed by the MOO algorithm to solve the MOO problem.
-The columns named HV and DU contain the values of the HV and DU indicators.
-The results of the numerical experiments are divided with respect to the three algorithms used: GSA is described in Section \ref{sec:03}, MGSA is discussed in Section \ref{sec:03} (item 2), SGSA is presented in Section \ref{sec:04} (item 3).
-
-The experimental results show that by reusing search information when optimizing 25 functions from the set $\Phi_p (y)$ the total computational volume can be reduced  by more than 9.5 times without resorting to any additional computational resources.
+The experimental results show that by reusing search information when optimizing 25 functions from the set $\Phi_p (y)$ the total computational iterations can be reduced  by more than 9.5 times without resorting to any additional computational resources.
 At the same time, the MGSA algorithm performs the least number of iterations (the number of calculations of objective function values), while the SGSA algorithm provides a better approximation of the Pareto domain both in terms of approximation completeness and coverage uniformity.
@@
 \hline
-\multirow{2}{*}{\begin{tabular}[c]{@{}c@{}}Number of \\ functions in $\Phi_p (y)$\end{tabular}} & \multicolumn{3}{c}{Average number of   iterations} \\
+\multirow{2}{*}{\begin{tabular}[c]{@{}c@{}}Number of \\ functions in $\Phi_p (y)$\end{tabular}} & \multicolumn{3}{c}{\begin{tabular}[c]{@{}c@{}}Average number \\ of iterations\end{tabular}} \\
                                                                                         & GSA                           & MGSA                         & SGSA                         \\ \hline
@@
 The objective functions of the MOO problems were determined, as before, using the GKLS generator \cite{c43}.
-When solving the problems, the accuracy $\varepsilon=0.05$ and the reliability of the method $r=5.6$ were used.
+When solving the problems, the accuracy $\varepsilon=0.05$ and the reliability $r=5.6$ were used.
 The results of numerical experiments are presented in Table \ref{tab:4}.
@@
 \centering
-\caption{The results of a series of experiments to solve five-dimensional bi-criteria MOO problems}
+\caption{The results of a series of experiments to solve bi-criteria five-dimensional MOO problems}
 \label{tab:4}
@@
 
-This paper proposes an effective method for solving computationally intensive multicriteria optimization problems which require computing several Pareto-optimal decisions.
-It is also assumed that efficiency criteria may be multiextremal, and computation of function values may require a large amount of calculations.
-The proposed approach is based on the reduction of multicriteria optimization problems to one-dimensional global optimization problems that can be solved using effective statistical information algorithms of global search.
+This paper proposes an efficient method for solving computationally intensive multi-objective optimization problems which require computing several Pareto-optimal decisions.
+It is also assumed that objective functions may be multiextremal, and computation of function values may require a large amount of calculations.
+The proposed approach is based on the reduction of multi-objective optimization problems to one-dimensional global optimization problems that can be solved using efficient information-statistical algorithms of global search.
 The novelty of the proposed approach consists in the possibility of solving several global optimization problems simultaneously, which allows multiple Pareto-optimal decisions to be obtained.
-Besides, such approach provides for reuse of the computed search information, which considerably reduces computational effort for solving multicriteria optimization problems.
+Besides, such approach provides for reuse of the computed search information, which considerably reduces computational effort for solving multi-objective optimization problems.
 
-The results of numerical experiments show that by using the developed approach it is possible to achieve a significant reduction of computational complexity when solving multicriteria optimization problems.
+The results of numerical experiments show that by using the developed approach it is possible to achieve a significant reduction of computational complexity when solving multi-objective optimization problems.
 
-It should be noted in conclusion that the proposed approach is quite promising and requires further research in this area.
-First of all, it is necessary to continue numerical experiments on solving multicriteria optimization problems with a greater number of partial efficiency criteria and for higher dimensionality of optimization problems to be solved.
+It should be noted in conclusion that the proposed approach is promising and requires further research in this area.
+First of all, it is necessary to continue numerical experiments on solving multi-objective optimization problems with a greater number of objective functions and for higher dimensionality of optimization problems to be solved.
 The possibility of organizing parallel computations for high-performance systems should also be evaluated.

### 2021-02-13 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
Before translation

@@
 \usepackage{ljm}
+
+\usepackage{makecell}
 
@@
 \begin{abstract} % You shouldn't use formulas and citations in the abstract.
-%абстракта пока нет, подготовим позже
+В работе рассматриваются mixed-integer global optimization problems.
+Обсуждается параллельный алгоритм для решения задач данного класса, основанный на information-statistical approach к решению задач непрерывной глобальной оптимизации.
+Для рассматриваемого параллельного алгоритма предложена схема локальной настройки, основанная на предположении слабой многоэкстремальности задачи.
+Приведены результаты сравнения последовательной версии алгоритма с другими методами аналогичного назначения. 
+Эффективность распараллеливания алгоритма подтверждена решением серии mixed-integer global optimization problems на Lobachevsky supercomputer. 
 \end{abstract}
@@
 \begin{eqnarray}\label{lR}
-& h_i = \left[\frac{\sqrt{(z_i-z_\nu^*)(z_{i-1}-z_\nu^*)}}{\mu_\nu}+1.5^{-q}\right]^{-1}, \; \nu(x_{i-1})=\nu(x_{i}),\nonumber\\
-& h_i = \left[\frac{z_i-z_\nu^*}{\mu_\nu}+1.5^{-q}\right]^{-1}, \; \nu(x_{i-1})<\nu(x_{i}),\\
-& h_i = \left[\frac{z_{i-1}-z_\nu^*}{\mu_\nu}+1.5^{-q}\right]^{-1}, \; \nu(x_{i-1})>\nu(x_{i}).\nonumber
+& h_i = \left[\frac{\sqrt{(z_i-z_\nu^*)(z_{i-1}-z_\nu^*)}}{\mu_\nu}+10^{-q}\right]^{-1}, \; \nu(x_{i-1})=\nu(x_{i}),\nonumber\\
+& h_i = \left[\frac{z_i-z_\nu^*}{\mu_\nu}+10^{-q}\right]^{-1}, \; \nu(x_{i-1})<\nu(x_{i}),\\
+& h_i = \left[\frac{z_{i-1}-z_\nu^*}{\mu_\nu}+10^{-q}\right]^{-1}, \; \nu(x_{i-1})>\nu(x_{i}).\nonumber
 \end{eqnarray}
-Здесь целочисленный неотрицательный параметр $q$ характеризует «степень локальности» поиска: при $q=0$ поиск будет иметь ``глобальный'', а при $q>>1$ -- ``локальный'' характер. 
+Здесь целочисленный неотрицательный параметр $q$ характеризует «степень локальности» поиска: при $q=0$ поиск будет иметь ``глобальный'', а при $q>1$ -- ``локальный'' характер. 
 В самом деле, множитель $h_i$ будет принимать наибольшее значение равное $1.5^q$ лишь для тех интервалов, граничные точки которых соответствуют текущему минимуму, т.е.
@@
 
-%Ниже пока переводить не надо
+%Переводить до конца, т.е. до acknowledgments
 
-Первая серия экспериментов была проведена с использованием последовательной версии предложенного алгоритма с целью его сравнения с известными методами аналогичного назначения.
+Первая серия экспериментов была проведена с использованием последовательной версии алгоритма без использования local tuning с целью его сравнения с известными методами аналогичного назначения.
 
 Предложенный в данной работе алгоритм PMGAL (в его последовательном варианте) сравнивался с genetic algorithm for solving mixed-integer optimization problems, реализованным в Matlab Global Optimization Toolbox.
-В Table \ref{tab:1} приведено число испытаний, потребовавшееся для решения известных тестовых mixed-integer problems данными методами.
+В Table \ref{tab:1} приведено число trials, потребовавшееся для решения данными методами ряда тестовых mixed-integer problems, взятых из \cite{Deep,Floudas}.
 Для обоих методов была использована одинаковая точность поиска решения $10^{-2}$.
@@
 		\hline
-		%\noalign{\smallskip}		
 	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c|}{ GA } &  \multicolumn{2}{c|}{MIGSA} \\
-		%\noalign{\smallskip} 
 		\cline{2-3} \cline{4-5} 
-		%\noalign{\smallskip}
 		 & $k$ & $t$ &  $k$ & $t$  \\
-		%\noalign{\smallskip} 
 		\hline 
-		%\noalign{\smallskip}
 		 Problem 2 \cite{Floudas}&	481 &	0.0601 & 	417 &	0.04 \\
@@
 		 Problem 7 \cite{Deep}   &	841 &	0.0736 &  372	 &	0.017 \\
-		%\noalign{\smallskip}
 		\hline
@@
 
-%Последующие серии экспериментов были выполнены для оценки ускорения параллельной версии предложенного алгоритма MIPIA с использованием как CPU, так и GPU.
-В данных экспериментах решалась серия из 100 тестовых mixed-integer problems, сгенерированных случайных образом.
-
-Известным генератором тестовых задач непрерывной многоэкстремальной оптимизации является GKLS-генератор \cite{Gaviano}.
-Он позволяет генерировать тестовые функции заданной размерности с заданными свойствами: известным числом локальных минимумов, известной областью притяжения глобального решения, и т.д.
+Следующий эксперимент был проведен для оценки влияния local tuning на ускорение сходимости алгоритма.
+В рамках данного эксперимента решалась серия задач, для построения которых использовался генератор GKLS \cite{Gaviano} для тестовых задач непрерывной многоэкстремальной оптимизации.
+Генератор GKLS позволяет порождать непрерывные тестовые функции заданной размерности с заданными свойствами: известным числом локальных минимумов, известной областью притяжения глобального решения, и т.д.
 %This generator of multiextremal functions is often used for the investigations of the global optimization algorithms \cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
@@
 In Fig.~\ref{example} (a) and (b), the contour plots of two-dimensional GKLS functions are presented.
-Figures also shows the points of the trials performed by the method until the required accuracy $\epsilon=10^{-2}$ was achieved.
+Figures also shows the points of the trials performed by the global search until the required accuracy $\epsilon=10^{-2}$ was achieved.
 
@@
 
-В проведенных экспериментах GKLS-генератор использовался как основа для построения mixed-integer problems. Правила, позволяющие генерировать тестовые задачи подобного вида, заключаются в следующем.
+В проведенных экспериментах генератор GKLS использовался как основа для построения mixed-integer problems. Правила, позволяющие генерировать тестовые задачи подобного вида, заключаются в следующем.
 
 \begin{enumerate}
-	\item С помощью генератора GKLS порождается непрерывная многоэкстремальная функция $\psi(y), \; y\in D, \; D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\}$.
+	\item С помощью генератора порождается непрерывная многоэкстремальная функция $\psi(y), \; y\in D, \; D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\}$.
 Глобальный минимум данной функции достигается в известной точке $y'=(y'_1,...,y'_N)$ и равен $\psi'=\psi(y')=-1$.
@@
 	\[
-	C = 4 - \min_{y,u} \left\{ \phi(y,u) \right\}, \; y\in D, u \in U.
+	C = 4 - \max_{y,u} \left\{ \phi(y,u) \right\}, \; y\in D, u \in U.
 	\]
-	Минимум вогнутой функции $-\phi(y,u)$ достигается в одной из угловых точек области поиска, поэтому при размерности задачи порядка 10 его вычисление может быть выполнено методом перебора.
+	Максимум выпуклой функции $\phi(y,u)$ достигается в одной из угловых точек области поиска, поэтому при размерности задачи порядка 10 его вычисление может быть выполнено методом перебора.
 	\item Формируется многоэкстремальная mixed-integer function 
 	\[
-	\varphi(y,u) = \left(\psi(y) + \sum_{j=1}^M{u_j}\right)\left(C - \phi(y,u)\right).
+	\varphi(y,u) = \left(\psi(y) - \sum_{j=1}^M{u_j}\right)\left(C - \phi(y,u)\right).
 	\]
@@
 
-In the problems generated in our experiments, there were 5 discrete and 6 continuous parameters 
+In the problems generated in our experiments, there were 5 discrete and 4 continuous parameters 
 	\begin{eqnarray*}
-	& y\in D = \left\{ -1 \leq y_j\leq 1, 1\leq j \leq 6 \right\} \subset R^6, \\
+	& y\in D = \left\{ -1 \leq y_j\leq 1, 1\leq j \leq 4 \right\} \subset R^4, \\
 	& u\in U = \left\{ u_j \in  \left\{-1, -1/3, 0, 1/3, 1 \right\}, 1\leq j \leq 5 \right\}.
@@
 
+Было сгенерировано и решено 100 9-и мерных mixed-integer problems данного типа.
+С целью имитации вычислительной трудоемкости, присущей прикладным задачам оптимизации, расчет функции во всех проводимых экспериментах был усложнен дополнительными вычислениями, не меняющими вид функции и расположение ее минимумов.
+The accuracy of the search was equal to $10^{-2}$.
+Computational experiments were carried out on Lobachevsky supercomputer.
+The node of supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.
+
+В Table \ref{tab:2} приведены результаты экспериментов, полученные для последовательной версии алгоритма PMGAL с использованием различных стратегий чередования локального и глобального поиска.
+Для локального поиска в формулах (\ref{lR}) использовался параметр $q=3$, для глобального -- $q=0$.
+В первом столбце таблицы отражена частота чередования указанных значений.
+Например, 1:1 соответствует чередованию локального и глобального поиска на каждой итерации, 1:2 соответствует одному шагу локального поиска и двум шагам глобального поиска, и т.д.
+Во втором столбце таблицы отражено среднее число итераций $K_{av}$ (тыс.), требуемое для решения задачи с заданной точностью.
+
+\begin{table}
+	\caption{Сравнение стратегий local tuning}
+	\label{tab:2}
+	\center
+	\begin{tabular}{|c|c|}
+		\hline		
+		\textit{loc} : \textit{glob} & $K_{av}$   \\
+		\hline 
+		1 : 1 & 102.6\\
+		1 : 2 & 89.2\\
+		1 : 4 & 97.5\\
+		1 : 8 & 127.1\\
+		1 : 16 & 144.2\\
+		\hline
+	\end{tabular}
+\end{table}
+
+Результаты экспериментов показывают, что наименьшее число итераций методу с локальной настройкой требуется при использовании стратегии чередования 1:2, т.е.
+один шаг локального уточнения чередуется с двумя шагами глобального поиска.
+Будем использовать эту стратегию и в дальнейших экспериментах.
+
+\begin{table}
+	\caption{Среднее время работы $T_{av}$ параллельного алгоритма с локальной настройкой}
+	\label{tab:3}
+	\center
+	\begin{tabular}{|c|c|c|c|}
+		\hline	
+	\diaghead{\theadfont Diag Column} { $N_{p}$}{$N_{t}$} & 1 & 4 & 8\\
+
+		\hline		
+		2  & 199.0 & 65.4 & 61.0  \\
+		5  & 91.2 & 28.0 &  26.1 \\
+		17  & 37.2 & 10.8 & 7.5 \\
+		65  & 29.3 & 8.7 & 5.4 \\
+		\hline
+	\end{tabular}
+\end{table}
+
+Последняя серия экспериментов была выполнена для оценки ускорения параллельной версии алгоритма PMGAL. 
+В Table \ref{tab:3} приведено среднее время решения $T_{av}$ (в секундах) в зависимости от задействованного числа процессов $N_{p}$ и потоков $N_{t}$ в каждом процессе.
+При проведении экспериментов было задействовано от 1 до 16 узлов кластера в зависимости от числа параллельных процессов.
+Ускорение, которое демонстрирует параллельный алгоритм, не является идеальным, но -- вполне приемлемое.
+
+\section{Conclusions}
+
+В данной статье представлены результаты, связанные с разработкой параллельных алгоритмов глобальной оптимизации для решения многоэкстремальных задач, в которых часть параметров является непрерывными, а часть -- дискретными.
+Рассматриваемые задачи могут включать в себя также невыпуклые ограничения, учет которых организован с использованием специальной индексной схемы. 
+Для задач указанного класса предложен параллельный алгоритм их решения.
+Основной принцип организации параллельных вычислений в алгоритме основан на одновременном проведении нескольких испытаний в различных точках области поиска.
+
+В параллельном алгоритме глобальной оптимизации использована техника локальной настройки, основанная на предположении слабой многоэкстремальности задачи.
+Данное предположение является верным, в частности, на заключительной фазе поиска, когда уже выполнено одно или несколько испытаний в окрестности глобального решения, и осталось найденную оценку решения уточнить.
+
+В последовательном варианте предложенный алгоритм не уступает методам аналогичного назначения, реализованным в Matlab Global Optimization Toolbox.
+В параллельном варианте предложенный алгоритм демонстрирует приемлемую масштабируемость вплоть до ипользования десятков параллельных процессов.
+
 \begin{acknowledgments}
-%We are grateful to C.~C.~Surname3 and the reviewers for careful reading of the manuscript and helpful remarks.
 This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.

### 2021-02-13 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex

@@
 It is also assumed that efficiency criteria may be multiextremal, and the cost of calculating function values may be quite high.
-The proposed approach is based on the reduction of multi-objective optimization problems to one-dimensional global optimization problems that can be solved using effective information-statistical algorithms of global search.
+The proposed approach is based on the reduction of multi-objective optimization problems to one-dimensional global optimization problems that can be solved using efficient information-statistical algorithms of global search.
 One key innovation of the developed approach consists in the possibility of solving several global optimization problems simultaneously, which allows multiple Pareto-optimal decisions to be obtained.
@@
 In the proposed approach, we use for scalarization the compromise programming method \cite{c2,c11}, where solution of the MOO problem consists in finding the efficient decision corresponding most closely to the values of the objective functions of the specified reference decision $y^0 \in D$.
-In this case, a possible statement of the statement scalar optimization problem can have the form
+In this case, a possible statement of scalar optimization problem can have the form
 \begin{equation}
@@
 Obtaining different efficient decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multi-extremal functions $F(\lambda, y)$ from (\ref{eq:04}).
-These functions $F(\lambda, y)$ for different values of the coefficients $\lambda \in \Lambda$ can be optimized sequentially.
+These functions $F(\lambda, y)$ can be optimized sequentially.
 This determines a multi-stage scheme for solving the MOO problem, when at each stage DM specifies the necessary scalarization coefficients, then the resulting scalar optimization problem is solved, after which DM analyzes the efficient decisions found.

### 2021-02-15 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
Many corrections

@@
 Victor Gergel\orcidID{0000-0002-4013-2329} \and
-Vladimir Grishagin \and
+Vladimir Grishagin\orcidID{0000-0002-2884-3670} \and
 Evgeniy Kozinov\orcidID{0000-0001-6776-0096} 
@@
 Choosing the optimal options (decision making) in situations with many different alternatives is a problem that occurs in almost every domain of human activity.
-In the simplest cases, decision making problems can be viewed as optimization problems of various kinds such as convex programming, discrete optimization, nonlinear programming, etc.
+In many cases, decision making problems can be viewed as optimization problems of various kinds such as convex programming, discrete optimization, nonlinear programming, etc.
 In more complex situations, the objective functions that determine the effectiveness of decisions can be multiextremal, and decision making will require solving global optimization problems.
@@
 One of the approaches most commonly used in the search for efficient decisions is scalarization of the vector efficiency criterion, when a MOO problem is reduced to solving one or several scalar (in the general case, global) optimization problems in which the objective functions are converted to the integrated scalar function.
-The family of optimization problems generated in such approach is further referred to as multiple global optimization problem (MGO).
-Within this approach, we can distinguish lexicographic optimization methods where objective functions are ordered by importance, thus  allowing the optimization of the functions to be carried out sequentially as their importance decreases [10].
+The family of optimization problems generated in such approach is further referred to as multiple global optimization (MGO) problem.
+Within this approach, we can distinguish lexicographic optimization methods where objective functions are ordered by importance, thus  allowing the optimization of the functions to be carried out sequentially as their importance decreases \cite{c10}.
 Possible scalarization methods also include various methods of the efficiency criteria convolution such as the weighted sum method, the compromise programming method, the reference point method, the weighted min-max method, etc.
@@
 \end{equation}
-where $f(y) = (f_1(y), f_2(y), \dots, f_s(y))$ are objective functions (efficiency criteria), $y = (y_1, y_2, \dots , y_N)$ is the vector of varied parameters, and a $N$ is the dimensionality of the multi-objective optimization problem to be solved.
+where $f(y) = (f_1(y), f_2(y), \dots, f_s(y))$ are objective functions (efficiency criteria), $y = (y_1, y_2, \dots , y_N)$ is the vector of varied parameters, and $N$ is the dimensionality of the multi-objective optimization problem to be solved.
 The set of possible parameter values (search domain) $D$ is usually an $N$-dimensional hypercube
@@
 \label{eq:04}
-\min {\varphi(y)} = F(\lambda, y), y \in D,
+\min_{y \in D} {\varphi(y)} = \min_{y \in D} F(\lambda, y),
 \end{equation}
@@
 \label{eq:05}
-\min F(\lambda,y)=\min 1/s \sum_{i=1}^s {\lambda_i (f_i (y)-f_i (y^0 ))^2}, y \in D,
+\min_{y \in D} F(\lambda,y)=\min_{y \in D} \left\{ \frac{1}{s} \sum_{i=1}^s {\lambda_i (f_i (y)-f_i (y^0 ))^2} \right\},
 \end{equation}
-where $F(\lambda,y)$ is the standard deviation of the values of the objective functions $f_i$, $1 \leq i \leq s$, for the decision $y \in D$ and for the specified reference decision $y^0 \in D$, while the coefficients $0 \leq \lambda_i \leq 1$, $1 \leq i < s$, are measures of the significance of the approximation accuracy for each variable $y_i$, $1 \leq i \leq N$, separately.
+where $F(\lambda,y)$ is the standard deviation of the values of the objective functions $f_i$, $1 \leq i \leq s$, for the decision $y \in D$ and for the specified reference decision $y^0 \in D$, while the coefficients $\lambda_i$, $0 \leq \lambda_i \leq 1$, $1 \leq i \leq s$, are measures of the significance of the approximation accuracy for each variable $y_i$, $1 \leq i \leq N$, separately.
 Without loss of generality, we can assume that the area of possible values of the coefficients $\lambda$ is a set
@@
 \label{eq:08}
-|F(\lambda, y') - F(\lambda, y'')| \leq L\|y' - y''\|.
+|F(\lambda, y') - F(\lambda, y'')| \leq L\|y' - y''\|, \; y',y''\in D.
 \end{equation}
@@
 \label{eq:09}
-\min \varphi (y(x))= \min \varphi(y), x \in [0,1], y \in D.
+\min_{x \in [0,1]} \varphi (y(x))= \min_{y \in D} \varphi(y).
 \end{equation}
@@
 \label{eq:10}
-|\varphi (y(x'))- \varphi (y(x''))| \leq H |x'-x''|^{1/N},
+|\varphi (y(x'))- \varphi (y(x''))| \leq H |x'-x''|^{1/N}, \; x',x''\in[0,1],
 \end{equation}
@@
 \textbf{3.
-Solving the one-dimensional reduced optimization problem.} Using the dimensionality reduction results in one more additional advantage of the proposed approach: many well-known one-dimensional global search algorithms (possibly, after some additional generalisation) can be used to solve the initial multidimensional MOO problem from (\ref{eq:01}) -- see e.g.
-\cite{c22,c23,c24,c25,c26,c27,c28,c29}.
+Solving the one-dimensional reduced optimization problem.} Using the dimensionality reduction results in one more additional advantage of the proposed approach: many well-known one-dimensional global search algorithms (possibly, after some additional generalisation) can be used to solve the initial multidimensional MOO problem from (\ref{eq:01}) \cite{c22,c23,c24,c25,c26,c27,c28,c29}.
 At the same time, it should be noted that most works where dimensionality reduction is a key feature for solving multiextremal problems like (\ref{eq:04}) rely on the information-statistical theory of global search \cite{c15}.
-This theory has provided the a basis for developing a large number of efficient methods for multi-extremal optimization -- see, e.g., \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
+This theory has provided the a basis for developing a large number of efficient methods for multi-extremal optimization \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
 
@@
 
-This estimate is determined by means of characteristics of intervals $R(i)$, $1 \leq i<k$, whose values should be proportional to the degree of possibility that the global minimum is located  in these intervals.
+This estimate is determined by means of characteristics $R(i)$ of intervals $(x_{i-1},x_i)$, $1 < i \leq k$, whose values should be proportional to the degree of possibility that the global minimum is located  in these intervals.
 The type of these characteristics depends on the global optimization algorithm used -- thus, for example, for an algorithm which constructs a uniform dense grid in the global search domain, the characteristic is simply the interval length
@@
 \label{eq:12}
-R(i)=(x_i-x_{i-1}), 1 \leq i<k.
+R(i)=(x_i-x_{i-1}), 1 < i \leq k.
 \end{equation}
 
-For the algorithms proposed in \cite{c22,c23}, when dimensionality reduction is applied, the characteristic is an estimate of the minimum possible value of the function $\varphi(y(x))$ to be minimized on the interval $(x_{i-1},x_i )$, $1 \leq i < k$, i.e.
+For the algorithms proposed in \cite{c22,c23}, when dimensionality reduction is applied, the characteristic is an estimate of the minimum possible value of the function $\varphi(y(x))$ to be minimized on the interval $(x_{i-1},x_i )$, $1 < i \leq k$, i.e.
 \begin{equation}
 \label{eq:13}
-R(i) = 0.5 H \varrho_i - 0.5 (z_{i-1} + z_i ), \varrho_i = (x_i-x_{i-1} ) ^ {1⁄N}, 1 \leq i < k,
+R(i) = 0.5 H \rho_i - 0.5 (z_{i-1} + z_i ), \rho_i = (x_i-x_{i-1} ) ^ {1⁄N}, 1 < i \leq k,
 \end{equation}
-where $H$ is the H{\" o}lder constant from (\ref{eq:10}) for the reduced global optimization problem being solved (\ref{eq:09}), $z_i = \varphi(y(z_i))$, $1 \leq i < k$ and $N$ is the dimensionality of the problem from (\ref{eq:01})).
+where $H$ is the H{\" o}lder constant from (\ref{eq:10}) for the reduced global optimization problem being solved (\ref{eq:09}), $z_i = \varphi(y(z_i))$, $1 \leq i \leq k,$ and $N$ is the dimensionality of the problem from (\ref{eq:01})).
 For the global search algorithm (GSA) \cite{c15,c24} developed in the framework of the information-statistical approach, the characteristic is
@@
 \label{eq:14}
-R(i)=m\varrho_i+\frac{z_i-z_{i-1}^2}{m\varrho_i}-2(z_{i-1}+z_i ),\varrho_i=(x_i-x_{i-1})^{1/N}, 1 \leq i < k,
+R(i)=m\rho_i+\frac{z_i-z_{i-1}^2}{m\rho_i}-2(z_{i-1}+z_i ),\rho_i=(x_i-x_{i-1})^{1/N}, 1 < i \leq k,
 \end{equation}
@@
 \label{eq:15}
-m= r M, M = \max |z_i-z_{i-1} | ⁄ \varrho_i, 1 \leq i<k
+m= r M, M = \max \{ |z_i-z_{i-1} | ⁄ \rho_i, 1 < i \leq  k \}
 \end{equation}
@@
 
-\textit{Step 1.} Calculate characteristics of the  intervals $R(i)$, $1 \leq i < k$, and determine the interval with the maximum characteristic
+\textit{Step 1.} Calculate characteristics of the  intervals $R(i)$, $1 < i \leq k$, and determine the interval with the maximum characteristic
 \begin{equation}
 \label{eq:16}
-R(t) = \max R(i), 1\leq i < k.
+R(t) = \max \{ R(i), 1< i \leq k \}.
 \end{equation}
@@
 \label{eq:18}
-\varrho_t \leq \varepsilon,
+\rho_t \leq \varepsilon,
 \end{equation}
-where $\varrho_t = (x_t-x_{t-1})^{1⁄N}$, $t$ from (\ref{eq:16}) and $\varepsilon > 0$ is the specified accuracy of the solution.
+where $\rho_t = (x_t-x_{t-1})^{1⁄N}$, $t$ is from (\ref{eq:16}) and $\varepsilon > 0$ is the specified accuracy of the solution.
 If the stopping condition (\ref{eq:18}) is met, then the solving of the optimization problem is stopped, otherwise $k=k+1$ is assumed and the next global search iteration begins.
@@
 \label{eq:19}
-z_k^* = \min z_i, 1 \leq i<k.
+z_k^* = \min \{ z_i, 1 \leq i \leq \}.
 \end{equation}
@@
 It should be noted again that the computational scheme discussed above is quite general.
-Many global search algorithms can be represented within this scheme, as evidenced, in particular, by examples (\ref{eq:12})-(\ref{eq:14}) and other global optimization algorithms -- see, for example, \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
+Many global search algorithms can be represented within this scheme, as evidenced, in particular, by examples (\ref{eq:12})--(\ref{eq:14}) and other global optimization algorithms -- see, for example, \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
 
-Convergence conditions of the algorithms formulated in the form (\ref{eq:16})-(\ref{eq:18}).
-One of sufficient conditions for convergence of algorithms is, for example, the requirement that the characteristic of the interval containing the global minimum point should take on the maximum value at step 1 of the scheme (\ref{eq:16})-(\ref{eq:18}) during the global search iterations.
+Convergence conditions of the algorithms formulated in the form (\ref{eq:16})--(\ref{eq:18}).
+One of sufficient conditions for convergence of algorithms is, for example, the requirement that the characteristic of the interval containing the global minimum point should take on the maximum value at step 1 of the scheme (\ref{eq:16})--(\ref{eq:18}) during the global search iterations.
 This condition is satisfied, for example, for the multidimensional generalized algorithms proposed in \cite{c22,c23} when the H{\"o}lder constant from (\ref{eq:10}) is specified exactly.
@@
 \label{eq:21}
-\Phi_p (y)=\{ F(\lambda_1,y), F(\lambda_2,y),\dots,F(\lambda_p,y)  \},  \lambda_i \in \Lambda, 1 \leq i \leq p
+\Phi_p (y)=\{ F(\lambda_1,y), F(\lambda_2,y),\dots,F(\lambda_p,y)  \},  \lambda_i \in \Lambda, 1 \leq i \leq p,
 \end{equation}
@@
 \label{eq:22}
-A_k=\{(x_i,z_i,f_i )^T:1 \leq i \leq k\},
+A_k=\{(x_i,z_i,f_i )^T:1 \leq i \leq k\}
 \end{equation}
-obtained in the calculations (see (\ref{eq:14})-(\ref{eq:17})).
-In (\ref{eq:22}) $x_i$, $1 \leq i \leq k$,  are the reduced points of performed global search iterations ordered in ascending order of coordinates,  $z_i$, $f_i$, $1 \leq i \leq k$, are values of the scalar function $F(\lambda, y)$, from (\ref{eq:04}) and objective functions $f_i(y)$, $1 \leq i \leq s$, from (\ref{eq:01}) of the current optimization problem to be solved at the points $x_i=y(x_i)$, $1 \leq i \leq k$.
+obtained in the calculations (see (\ref{eq:14})--(\ref{eq:17})).
+In (\ref{eq:22}) $x_i$, $1 \leq i \leq k$,  are the reduced points of performed global search iterations ordered in ascending order of coordinates,  $z_i$, $f_i$, $1 \leq i \leq k$, are values of the scalar function $F(\lambda, y)$, from (\ref{eq:04}) and objective functions $f(y)$ from (\ref{eq:01}) of the current optimization problem to be solved at the points $x_i=y(x_i)$, $1 \leq i \leq k$.
 By using search information $A_k$ from (\ref{eq:22}) when choosing next search iterations it is possible to solve global optimization problems more efficiently and to provide convergence of algorithms only to the global minima of multiextremal functions being minimized.
 
-It is important to note that since the set $\Phi_p(y)$ of functions being optimized  simultaneously is generated from the same MOO problem from (\ref{eq:01}), the existence of the set $A_k$ from (\ref{eq:12}) allows us to adjust the results of all previously performed calculations of the values of the objective functions $f_i(y)$, $1 \leq i \leq s$, to the values of the next optimized function $F(\lambda, y)$ from (\ref{eq:04}) without repeating any time consuming calculations of values, i.e.
+It is important to note that since the set $\Phi_p(y)$ of functions being optimized  simultaneously is generated from the same MOO problem from (\ref{eq:01}), the existence of the set $A_k$ from (\ref{eq:22}) allows us to adjust the results of all previously performed calculations of the values of the objective functions $f(y)$ to the values of the next optimized function $F(\lambda, y)$ from (\ref{eq:04}) without repeating any time consuming calculations of values, i.e.
 \begin{equation}
@@
 Thus, all the search information $A_k$ from (\ref{eq:22}), recalculated according to (\ref{eq:23}), can be reused to continue solving the problems of the set $\Phi_p(y)$.
-Such a possibility provides a significant reduction in computations up to performing only a limited set of global search iterations (see the results of numerical experiments in Section \ref{sec:03}).
+Such a possibility provides a significant reduction in computations up to performing only a limited set of global search iterations (see the results of numerical experiments in Section \ref{sec:05}).
 
-     This type of information connectivity of functions in the set $\Phi_p(y)$ from (\ref{eq:21}) makes it possible to generalize the computational scheme (\ref{eq:16})-(\ref{eq:18}) for solving a single global optimization problem for the case of optimizing the functions of the set $\Phi_p(y)$ from (\ref{eq:21}) by adding a preliminary step of the search information  transformation.
+     This type of information connectivity of functions in the set $\Phi_p(y)$ from (\ref{eq:21}) makes it possible to generalize the computational scheme (\ref{eq:16})--(\ref{eq:18}) for solving a single global optimization problem for the case of optimizing the functions of the set $\Phi_p(y)$ from (\ref{eq:21}) by adding a preliminary step of the search information  transformation.
 
@@
 \label{eq:24}
-A_k=\{(x_i,\overrightarrow{z_i},f_i )^T : 1 \leq i \leq k\}
+A_k=\{(x_i,\overrightarrow{z_i},f_i )^T : 1 \leq i \leq k\},
 \end{equation}
@@
 \label{eq:25}
-\overrightarrow{z_i}=( z_i (1),z_i (2), \dots ,z_i (p)),z_i (j)=F(\lambda_j,y(x_i)), 1\leq i \leq k,1\leq j\leq p.
+\overrightarrow{z_i}=( z_i (1),z_i (2), \dots ,z_i (p)), \; z_i (j)=F(\lambda_j,y(x_i)), 1\leq i \leq k, 1\leq j\leq p.
 \end{equation}
 
-Accordingly, for each interval $(x_{i-1},x_i)$, $1 \leq i < k$, into which the segment [0,1] is divided, the following set of characteristics will be calculated:
+Accordingly, for each interval $(x_{i-1},x_i)$, $1 < i \leq k$, into which the segment [0,1] is divided, the following set of characteristics will be calculated:
 \begin{equation}
@@
 \label{eq:27}
-R_j(i)=m_j \varrho_i+\frac{(z_i(j)-z_{i-1}(j))^2}{m_j \varrho_i} - 2(z_{i-1}(j)+z_i(j)),1 \leq i < k,1 \leq j \leq p,
+R_j(i)=m_j \rho_i+\frac{(z_i(j)-z_{i-1}(j))^2}{m_j \rho_i} - 2(z_{i-1}(j)+z_i(j)),1 < i \leq k,1 \leq j \leq p,
 \end{equation}
@@
 \label{eq:28}
-\varrho_i=(x_i-x_{i-1} )^{1/N}  ,1 \leq i<k,
+\rho_i=(x_i-x_{i-1} )^{1/N}  ,1 \leq i<k,
 \end{equation}
@@
 \label{eq:29}
-m_j = r M_j, M_j=\frac{\max|z_i (j)-z_{i-1} (j)|}{\varrho_i} ,1\leq i < k, 1 \leq j \leq p.
+m_j = r M_j, M_j=\left\{ \frac{\max|z_i (j)-z_{i-1} (j)|}{\rho_i} ,1< i \leq k, 1 \leq j \leq \right\}.
 \end{equation}
-($m_j$, $1 \leq j \leq p$, is the estimate of the H{\"o}lder  constant in the condition (\ref{eq:10}) for the function $F(\lambda_j,y)$, $1 \leq j \leq p$, of the set $\Phi_p(y)$ from (\ref{eq:21})).
+($m_j$, $1 \leq j \leq p$, are the estimates of the H{\"o}lder  constant in the condition (\ref{eq:10}) for the functions $F(\lambda_j,y)$, $1 \leq j \leq p$, of the set $\Phi_p(y)$ from (\ref{eq:21})).
 
@@
 
-\textit{Step 1.} Compute characteristics of the intervals $R_j (i)$, $1 \leq i < k$, $1 \leq i \leq p$, and determine the function $F(\lambda_j,y)$, $1 \leq j \leq p$, whose search information contains the interval with the maximum characteristic
+\textit{Step 1.} Compute characteristics of the intervals $R_j (i)$, $1 < i \leq k$, $1 \leq i \leq p$, and determine the function $F(\lambda_j,y)$, $1 \leq j \leq p$, whose search information contains the interval with the maximum characteristic
 \begin{equation}
 \label{eq:30}
-R_q (t)=\max R_j (i), 1 \leq i < k, 1 \leq i \leq p.
+R_q (t)=\max \left\{ R_j (i), 1 < i \leq k, 1 \leq i \leq p \right\}.
 \end{equation}
@@
 \label{eq:32}
-\varrho_t \leq \varepsilon
+\rho_t \leq \varepsilon .
 \end{equation}
@@
 \label{eq:33}
-F'(\lambda_j,y)=(F(\lambda_j,y)-z_{min} (j))/H_j, 1 \leq j \leq p,
+F'(\lambda_j,y)=\frac{F(\lambda_j,y)-z_{min} (j)}{H_j}, 1 \leq j \leq p,
 \end{equation}
@@
 \label{eq:34_2}
-z_{min} (j)=\min_{y\in D}F(\lambda_j,y), 1 \leq j \leq p,
+z_{min} (j)=\min_{y\in D} \{ F(\lambda_j,y) \}, 1 \leq j \leq p,
 \end{equation}
-and $H_j$, $1 \leq j \leq p$, is the H{\"o}lder constant in the condition (\ref{eq:10}) for the function $F(\lambda_j,y)$, $1 \leq j \leq p$.
+and $H_j$, $1 \leq j \leq p$, are the H{\"o}lder constants for the functions $F(\lambda_j,y)$, $1 \leq j \leq p$.
 
@@
 
-Numerical experiments were performed on the Lobachevsky supercomputer (University of Nizhny Novgorod), (operating system -- CentOS 6.4, control system -- SLURM).
+Numerical experiments were performed on the Lobachevsky supercomputer of the University of Nizhni Novgorod (operating system -- CentOS 6.4, control system -- SLURM).
 One supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM.
@@
 a total of 16 CPU cores are available on the node).
-The numerical experiments were performed using the Globalizer system \cite{c36}.
+The numerical experiments were performed using the Globalizer system \cite{c38}.
 
@@
 The results of numerical experiments are presented in Table \ref{tab:2}.
-
-Table 2 shows in the first column the number of functions $\Phi_p (y)$ being optimized from (\ref{eq:21}).
+The table shows in the first column the number of functions $\Phi_p (y)$ from (\ref{eq:21}) being optimized.
 The columns named Iters contain the average number of iterations performed by the algorithm to solve the MOO problem.
@@
                                                                                         & Iters   & HV    & DU    &  & Iters   & HV     & DU    &  & Iters   & HV     & DU    \\ \hline
-5                                                                                       & 3771,8  & 6,455 & 0,25  &  & 1823,4  & 6,481  & 0,208 &  & 2090,5  & 6,479  & 0,217 \\
-10                                                                                      & 7941,6  & 6,485 & 0,183 &  & 1955,5  & 6,486  & 0,205 &  & 2224,4  & 6,486  & 0,191 \\
-25                                                                                      & 20456,2 & 6,504 & 0,143 &  & 2135,8  & 6,490  & 0,209 &  & 2456,1  & 6,493  & 0,188 \\ \hline
+5                                                                                       & 3771.8  & 6.455 & 0.25  &  & 1823.4  & 6.481  & 0.208 &  & 2090.5  & 6.479  & 0.217 \\
+10                                                                                      & 7941.6  & 6.485 & 0.183 &  & 1955.5  & 6.486  & 0.205 &  & 2224.4  & 6.486  & 0.191 \\
+25                                                                                      & 20456.2 & 6.504 & 0.143 &  & 2135.8  & 6.490  & 0.209 &  & 2456.1  & 6.493  & 0.188 \\ \hline
 \end{tabular}
@@
                                                                                         & GSA                           & MGSA                         & SGSA                         \\ \hline
-5                                                                                       & 754,4                         & 364,7                        & 418,1                        \\
-10                                                                                      & 794,2                         & 195,6                        & 222,4                        \\
-25                                                                                      & 818,2                         & 85,4                         & 98,2                         \\ \hline
+5                                                                                       & 754.4                         & 364.7                        & 418.1                        \\
+10                                                                                      & 794.2                         & 195.6                        & 222.4                        \\
+25                                                                                      & 818.2                         & 85.4                         & 98.2                         \\ \hline
 \end{tabular}
@@
                                                                                         & Iters   & HV    & DU    &  & Iters   & HV     & DU    &  & Iters   & HV     & DU    \\ \hline
-25                                                                                      & 1030039,6 & 2,440 & 0,543 & & 619755,6 & 2,441 & 0,681 & & 669743,8 & 2,364 & 0,668 \\ \hline
+25                                                                                      & 1030039.6 & 2.440 & 0.543 & & 619755.6 & 2.441 & 0.681 & & 669743.8 & 2.364 & 0.668 \\ \hline
 \end{tabular}
@@
 \bibitem{c22} Piyavskij, S.: An algorithm for finding the absolute extremum of a function, Computational Mathematics and Mathematical Physics \textbf{12}, 57--67 (1972).
-\bibitem{c23} Shubert, B.O.: A sequential method seeking the global maximum of a function. SIAM Journal on Numerical Analysis \textbf{9} , 379--388 (1972).
-\bibitem{c24} Strongin R.G.: Multiextremal Minimization. Autom. Remote Control. \textbf{7}, 1085--1088 (1970).
+\bibitem{c23} Shubert, B.O.: A sequential method seeking the global maximum of a function. SIAM Journal on Numerical Analysis \textbf{9}, 379--388 (1972).
+\bibitem{c24} Strongin R.G.: Multiextremal Minimization. Autom. Remote Control \textbf{7}, 1085--1088 (1970).
 \bibitem{c25} Galperin, E.A.: The cubic algorithm. Journal of Mathematical Analysis and Applications \textbf{112}, 635--640 (1985).
@@
 \bibitem{c27} Baritompa, W.: Accelerations for a variety of global optimization methods. Journal of Global Optimization \textbf{4}, 37--45 (1994).
-\bibitem{c28} Gergel, V.P.: A method of using derivatives in the minimization of multiextremum functions. Computational Mathematics and Mathematical Physics, \textbf{36}(6), 729--742 (1996).
+\bibitem{c28} Gergel, V.P.: A method of using derivatives in the minimization of multiextremum functions. Computational Mathematics and Mathematical Physics \textbf{36}(6), 729--742 (1996).
 \bibitem{c29} Sergeyev, Y.D.: Global one-dimensional optimization using smooth auxiliary functions. Mathematical Programming \textbf{81}, 127--146 (1998).
@@
 Communications in Nonlinear Science and Numerical Simulation, \textbf{23}, 328--342 (2015).
-\bibitem{c34} Grishagin, V., Israfilov, R., Sergeyev, Y.: Comparative efficiency of dimensionality reduction schemes in global optimization. AIP Conference Proceedings, 1776 (2016).
+\bibitem{c34} Grishagin, V., Israfilov, R., Sergeyev, Y.: Comparative efficiency of dimensionality reduction schemes in global optimization. AIP Conference Proceedings 1776 (2016).
 \bibitem{c35} Gergel, V.: An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems.
@@
 \doi{10.1063/1.4965367}
-\bibitem{c37} Gergel, V., Kozinov, E. Efficient multicriterial optimization based on intensive reuse of search information. Journal of Global Optimization, 2018, 71(1), 73--90.
+\bibitem{c37} Gergel, V., Kozinov, E. Efficient multicriterial optimization based on intensive reuse of search information. Journal of Global Optimization 71(1), 73--90 (2018).
 \bibitem{c38} Gergel, V., Barkalov, K., Sysoyev, A.
@@
 Autom.
-Remote Control, \textbf{75}(6), 1025--1040 (2014).
+Remote Control \textbf{75}(6), 1025--1040 (2014).
 \bibitem{c42} Bleuler, S., Laumanns, M., Thiele, L., Zitzler, E.: Pisa-a platform and programming language independent interface for search algorithms, Evolutionary Multi-Criterion Optimization.

### 2021-02-15 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
1. Таблица 4 дополнена

@@
                                                                                         & Iters   & HV    & DU    &  & Iters   & HV     & DU    &  & Iters   & HV     & DU    \\ \hline
-25                                                                                      & 1030039.6 & 2.440 & 0.543 & & 619755.6 & 2.441 & 0.681 & & 669743.8 & 2.364 & 0.668 \\ \hline
+25                                                                                      & 1030039.6 & 2.440 & 0.543 & & 619755.6 & 2.441 & 0.681 & & 669743.8 & 2.364 & 0.668 \\ 
+50                                                                                      & 2061634.2 & 2.502 & 0.572 & & 647508.2 & 2.533 & 0.691 & & 710418.6 & 2.417 & 0.680 \\ 
+100                                                                                     & 4104694.4 & 2.504 & 0.686 & & 683410.5 & 2.555 & 0.696 & & 728485.3 & 2.483 & 0.696 \\ \hline
 \end{tabular}
@@
 
-The results of our experiments show that with increasing dimensionality of the MOO problems to be solved, the trend that has been identified continues: the amount of computation (the number of global search iterations) is reduced by more than 1.6 times due to reuse of search information, the MGSA algorithm performs the least number of iterations, and the SGSA algorithm provides the best approximation of the Pareto domain.
+The results of our experiments show that with increasing dimensionality of the MOO problems to be solved, the trend that has been identified continues: the amount of computation (the number of global search iterations) is reduced by more than 5.6 times due to reuse of search information, the MGSA algorithm performs the least number of iterations, and the SGSA algorithm provides the best approximation of the Pareto domain.
 

### 2021-02-17 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
after translation, check up to Section 4

--- 2021-02-04 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
@@
 \begin{abstract} % You shouldn't use formulas and citations in the abstract.
-%абстракта пока нет, подготовим позже
+In this paper, we consider mixed-integer global optimization problems and propose a parallel algorithm for solving problems of this class based on information-statistical approach for solving continuous global optimization problems.
+Within this algorithm, we suggest using a local tuning scheme based on the assumption that the multiextremality of the discussed problem is weak.
+We also compare the sequential version of the algorithm with other similar methods.
+The effectiveness of parallelizing the algorithm has been confirmed by solving a series of mixed-integer global optimization problems on the Lobachevsky supercomputer.
 \end{abstract}
@@
 
-В работе рассматриваются global optimization problems и параллельные методы их решения.
-Методы решения задач указанного класса  обладают значительной вычислительной трудоемкостью, т.к.
-глобальный оптимум является интегральной характеристикой решаемой задачи и требует исследования всей области поиска.
-Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (как правило, неравномерного) в области изменения параметров.
-Особый интерес представляют задачи, в которых часть параметров может принимать лишь дискретные (например, целочисленные) значения (mixed-integer global optimization problems), т.к.
-для них сложнее построить оценки оптимума по сравнению с задачами, в которых все параметры являются непрерывными.
+Our research focuses on global optimization problems and parallel methods for their solution.
+Solving problems of this class requires substantial computational power since the global optimum as an integral characteristic of the problem requires the investigation of the entire search domain.
+Thus, the search for the global optimum is reduced to the construction of some coverage (as a rule, uneven) in the range of parameter change.
+This draws a particular interest to mixed-integer global optimization problems since it is more difficult to construct estimates of the optimum for such problems as compared to the continuous integer problems.
 
-Ситуация, когда некоторые из параметров характеризуются дискретностью, является широко распространенной в прикладных оптимизационных задачах.
-При этом дискретные параметры имеют, как правило, небольшое число значений и могут определять, например, марку используемого материала, вариант типовой компоновки деталей и т.п.
-Одновременно с этим целевая функция и ограничения обычно заданы не явно, а в виде некоторого алгоритма вычисления их значений в точках области поиска, т.е.
-представляют собой функции вида “black-box”.
-Процесс проверки ограничений и вычисления значения целевой функции в некоторой точке (называемый в дальнейшем trial) требует проведения численного моделирования и является трудоемкой операцией.
+So, it is quite typical when some of the parameters in the applied optimization problems are discrete.
+In this case, discrete parameters have, as a rule, a small number of values and can determine, for example, the grade of the material used, or the variation of the standard arrangement of parts, etc.
+At the same time, the objective function and constraints are usually specified not explicitly, but in the form of some algorithm for calculating their values at the points of the search domain, i.e.
+they represent black-box functions.
+The process of checking the constraints and calculating the value of the objective function at a certain point (hereinafter referred to as trial) requires numerical simulation and is a time-consuming operation.
 
-Методам решения mixed-integer problems посвящена обширная литература (см., например, обзоры \cite{Burer,Boukouvala}).
-Известные детерминированные методы решения задач данного класса основаны, как правило, на подходах Branch-and-Bound \cite{Belotti} or Branch-and-Reduce \cite{Vigerske}.
-Для решения таких задач могут применяться также nature-inspired algorithms \cite{Deep,Schluter}, которые так или иначе основаны на идеях случайного поиска.
+An extensive literature is devoted to methods for solving mixed-integer problems (see, for example, reviews \cite{Burer,Boukouvala}).
+Known deterministic methods for solving problems of this class are based, as a rule, on the Branch-and-Bound \cite{Belotti} or Branch-and-Reduce \cite{Vigerske}.
+.
+Such problems may also be solved by nature-inspired algorithms \cite{Deep,Schluter}, which in one way or another are based on a random search.
 
-Отметим ключевые моменты, специфичные для построения параллельных алгоритмов решения mixed-integer global optimization problems.
-Организовать распараллеливание процесса поиска можно несколькими способами.
-Во-первых, можно распараллеливать процесс проведения одного trial, т.е.
-вычисления функций задачи в одной точке области поиска.
-Данный путь будет давать ускорение, но является специфичным для каждой конкретной решаемой задачи.
-Во-вторых, можно распараллелить реализацию вычислительных правил алгоритма, обеспечивающих выбор точки очередного испытания.
-В этом случае способ распараллеливания будет зависеть от конкретного класса алгоритмов и, кроме того, часто эти правила достаточно просты и распараллеливать их нецелесообразно.
-В-третьих, можно организовать параллельное выполнение нескольких испытаний на одной итерации поиска.
-Этот подход и будет использоваться в данной работе, т.к.
-он характеризуется эффективностью (распараллеливается именно та часть вычислительного процесса, в которой выполняется основной объем вычислений) и общностью (применим для широкого класса алгоритмов глобальной оптимизации).
+Let us highlight the key points that are specific to the construction of parallel algorithms for solving mixed-integer global optimization problems.
+There are several ways to parallelize the search process.
+First, we can parallelize one trial, i.e.
+calculate functions of a problem at one point in the search domain.
+This approach will accelerate the calculations, but on the other hand, will have to be tailored for each particular problem.
+Second, we can parallelize the implementation of the computational rules of the algorithm that ensure the selection of a point for the next trial.
+In this case, the method of parallelization will depend on the specific class of algorithm.
+However, these rules are often quite simple, which makes their parallelization impractical.
+Third, it is possible to organize parallel execution of several trials during one search iteration.
+We will use this approach in this work due to its efficiency (it provides parallelization of the exact part of the computational process responsible for the bulk of the calculations) and generality (applicable to a wide class of global optimization algorithms).
 
-В данной работе нами предложен новый параллельный метод решения mixed-integer problems, основанный на information-statistical approach к решению задач глобальной оптимизации \cite{Strongin2000,Strongin2013}.
+In this paper, we propose a new parallel method for solving mixed-integer problems based on an information-statistical approach to solving global optimization problems \cite{Strongin2000,Strongin2013}.
 
-В рамках данного подхода 
+Within this approach 
 \begin{itemize}
 	\item 
-	решение многомерных задач сводится  к решению эквивалентных им одномерных, cоответствующая редукция основана на использовании space-filling curves;
+	solving multidimensional problems is reduced to solving equivalent one-dimensional problems, the corresponding reduction is based on the use of space-filling curves;
 	\item 
-	при решении constrained optimization problems каждое ограничение учитывается и обрабатывается отдельно, штрафные функции не используются;
+	when solving constrained optimization problems, each constraint is taken into account and processed separately, and penalty functions are not used;
 	\item
-	учитывается информация о неоднородном поведении функций задачи в разных подобластях областях области поиска, что позволяет ускорить процесс поиска решения за счет совмещения процедур локального уточнения текущего лучшего решения и его глобального обновления;
+	taking into account the information on the inhomogeneous behavior of the problem functions in different subdomains of the search domain makes it possible to speed up the process of finding a solution by combining the procedures for local refinement of the current best solution and its global update;
 	\item 
-	распараллеливание процесса поиска осуществляется за счет одновременного вычисления на каждой итерации нескольких значений целевой функции в разных точках search domain.	
+	parallelization of the search process is carried out by simultaneous calculation of several values of the objective function at different points of the search domain during each iteration.	
 \end{itemize}
 
-Текст статьи, отражающей результаты проведенного исследования, построен следующим образом. 
-В Section 2 дано краткое описание описание схемы редукции размерности с использованием space-filling curves, а также описана индексная схема учета ограничений.
-Здесь же приведена формулировка параллельного индексного алгоритма для решения задач непрерывной глобальной оптимизации.
-В Section 3 изложен способ учета локальных свойств функций задачи, позволяющий ускорить сходимость алгоритма.
+To provide a proper representation of the results of the study, we structured the text of the article as follows.
+Section 2 provides a brief description of the dimension reduction scheme using space-filling curves, as well as an index scheme for considering constraints.
+It also contains the formulation of a parallel index algorithm for solving continuous global optimization problems.
+Section 3 outlines a way to take into account the local properties of the problem functions to speed up the convergence of the algorithm.
+
 В Section 4 предложен подход к обобщению параллельного индексного алгоритма с целью решения mixed-integer problems.
 Приведен способ, позволяющий свести решение mixed-integer problem к решению информационно-связанного множества задач непрерывной оптимизации, которое может выполняться параллельно.
-Section 5 содержит результаты численных экспериментов.
-Здесь производится сравнение последовательной версии алгоритма с известными аналогами, а также показана эффективность распараллеливания алгоритма с локальной настройкой при решении серии multiextremal mixed-integer problem на суперкомпьютере.
+
+Section 5 contains the results of numerical experiments.
+There we compare the sequential version of the algorithm with known alternatives and also demonstrate the efficiency of parallelizing the algorithm with local tuning when solving the multiextremal mixed-integer problem series on a supercomputer.
 Section 6 concludes the paper.
@@
 
-Рассмотрим задачу глобальной оптимизации с непрерывными параметрами вида
+Let us consider a global optimization problem with continuous parameters of the form
 \begin{eqnarray}\label{problem}
@@
 \end{eqnarray}
-Будем предполагать, что objective function $\varphi(y)$ (hereafter denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
+We will assume that objective function $\varphi(y)$ (hereafter denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
 \[
@@
 
-Предположение о выполнении условия Липшица является типичным для многих подходов к построению параллельных алгоритмов глобальной оптмизации \cite{Evtushenko2009,Zilinskas2011}.
-При этом для решения многомерных задач оптимизации адаптируются алгоритмы решения одномерных задач, см., например, методы на основе разбиений области поиска \cite{Zilinskas2014,Sergeyev2017}.
+The assumption that the Lipschitz condition is satisfied is typical for many approaches to the construction of parallel algorithms for global optimization \cite{Evtushenko2009,Zilinskas2011}.
+In this case, the solution of a multidimensional optimization problem requires algorithms to be adjusted for solving one-dimensional problems, see, for example, methods based on partitions of the search domain \cite{Zilinskas2014,Sergeyev2017}.
 
-В данной работе мы будем использовать подход к сведению многомерных задач к одномерным на основе space-filling curves. 
-Используя отображения типа Peano-Hilbert curve $y(x)$, отображающие отрезок $[0,1]$ на $N$-dimensional domain (\ref{D}),
-можно свести многомерную задачу условной минимизации в области $D$ к одномерной задаче условной минимизации на отрезке $[0,1]$
+In this work, we will use an approach to reducing multidimensional problems to one-dimensional problems based on space-filling curves.
+Using mappings of the Peano-Hilbert curve $y(x)$ type, which project the segment $[0,1]$ onto the $N$-dimensional domain (\ref{D}),
+we can reduce the multidimensional conditional minimization problem in the domain $D$ to a one-dimensional conditional minimization problem in the segment $[0,1]$
 \begin{equation}\label{problem1}
@@
 
-Методы для построения чиcленных аппроксимаций Peano-Hilbert curve $y(x)$ (называемых \textit{evolvent}) рассмотрены в \cite{Sergeyev2013}.
-These evolvents are fractals generated by an iterative process, that fill in the hypercube $D$ with accuracy $2^{-M}$, where integer $M>0$ is the evolvent construction parameter.
-Примеры разверток с разными значениями $M$  для размерности $N=3$ приведены на Fig.~\ref{evolvents}.
+Methods for constructing numerical approximations of Peano-Hilbert curves $y(x)$ (called \textit{evolvents}) are discussed in \cite{Sergeyev2013}.
+These evolvents are fractals generated by an iterative process.
+They fill in the hypercube $D$ with accuracy $2^{-M}$, where integer $M>0$ is the evolvent construction parameter.
+Examples of evolvents with different values of $M$  for the dimensionality $N=3$ are shown in Fig.~\ref{evolvents}.
 
@@
 
-Задачу (\ref{problem1}) мы будем рассматривать в предположении, что функции $g_i(y(x))$ определены и вычислимы лишь на соответствующих множествах 
+We shall consider the problem (\ref{problem1}) under the assumption that the functions $g_i(y(x))$ are defined and computable only on the corresponding sets
 \[
@@
 
-Эти условия позволяют ввести классификацию точек $x \in [0,1]$ в соответствии с числом $\nu (x)$ ограничений, вычисленных в этой точке.
-The \textit{index} $\nu(x)$ также можно определить с помощью условий
+These conditions allow for the classification of points $x \in [0,1]$ in accordance with the number of constraints $\nu(x)$  calculated at this point.
+The \textit{index} $\nu(x)$ can also be determined using conditions
+
 \begin{equation}\label{nu}
@@
 \end{equation}
-где последнее неравенство исключается из рассмотрения, если $\nu=m+1$.
+where the last inequality is excluded from consideration if  $\nu=m+1$.
 
-Указанная схема редукции размерности сопоставляет многомерной задаче с липшицевой минимизируемой функцией и липшицевыми ограничениями одномерную задачу, в которой соответствующие функции удовлетворяют uniform H{\"o}lder condition (see \cite{Sergeyev2013}), i.e.,
+The indicated dimensionality reduction scheme associates a multidimensional problem with a Lipschitz objective function and Lipschitz constraints with a one-dimensional problem in which the corresponding functions satisfy the uniform  H{\"o}lder condition (see \cite{Sergeyev2013}), i.e.,
 \[
-\left|g_i(y(x_1))-g_i (y(x_2))\right| \leq H_i \left|x_1-x_2 \right|^{1/N}, \; x_1,x_2\in [0,1], \; 
-1\leq i \leq m+1,
+\left|g_i(y(x_1))-g_i (y(x_2))\right| \leq H_i \left|x_1-x_2 \right|^{1/N}, \; x_1,x_2\in [0,1], \; 1\leq i \leq m+1,
 \]
-где $N$ есть размерность исходной многомерной задачи, а H{\"o}lder coefficients $H_i$ связаны с Lipschitz constants $L_i$ исходной задачи соотношениями $H_i \leq 2L_i \sqrt{N+3}$.
+where $N$ is the dimensionality of the original multidimensional problem, and H{\"o}lder coefficients $H_i$ are related to Lipschitz constants $L_i$ of the original problem by the relation  $H_i \leq 2L_i \sqrt{N+3}$.
 
-Таким образом, \textit{a trial} в точке $x^k \in [0,1]$, выполненное на $k$-й итерации алгоритма, будет состоять из следующих действий:
+Thus, \textit{a trial} at a point $x^k \in [0,1]$ performed during the $k$-th iteration of the algorithm will feature the following steps:
 \begin{itemize}
-	\item определить \textit{image} $y^k=y(x^k)$ в соотвествии с отображением $y(x)$;
-	\item вычислить значения $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from (\ref{nu}). 
+	\item determine \textit{image} $y^k=y(x^k)$ in accordance with the mapping $y(x)$;
+	\item calculate values $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from (\ref{nu}). 
 \end{itemize}
-Пару значений 
+A value pair
 \begin{equation} \label{trial_result}
@@
 \end{equation}
-порождаемую испытанием в точке $x^k \in [0,1]$, будем называть \textit{trial outcome}.
+generated by the trial at a point $x^k \in [0,1]$, will be referred to as \textit{trial outcome}.
 
-Рассмотрим параллельный алгоритм для решения constrained global optimization problems, предложенный в работах \cite{Strongin2000,Strongin2013} и основанный на индексной схеме учета ограничений.
+Let us consider a parallel algorithm for solving constrained global optimization problems proposed in  \cite{Strongin2000,Strongin2013} and based on an index scheme for the accounting of constraints.
 
-При описании параллельного алгоритма мы будем предполагать, что в нашем распоряжении имеется $p$ вычислительных элементов (узлов, процессоров или ядер), которые будут использоваться для параллельного проведения $p$ trials одновременно.
+To describe a parallel algorithm, we will assume that we have $p$ computational elements (nodes, processors, or cores) at our disposal, which will be used for parallel execution of $p$ trials simultaneously.
 
-На первой итерации метода $p$ trials проводятся параллельно в произвольных различных точках $x^i\in(0,1)$, $1\leq i \leq p$.
-Пусть выполнено $n\geq 1$  итераций метода, в процессе которых были проведены испытания в $k=np$ точках $x^i, 1\leq i \leq k$.
-Тогда точки $x^{k+1},...,x^{k+p}$ поисковых испытаний $(n+1)$-ой итерации определяются в соответствии со следующими правилами.
+During the first iteration of the method, $p$ trials trials are carried out in parallel at different arbitrary points $x^i\in(0,1)$, $1\leq i \leq p$.
+Let $n\geq 1$ iterations of the method be performed with trials at $k=np$ points $x^i, 1\leq i \leq k$.
+Then points $x^{k+1},...,x^{k+p}$ of the search trials during the $(n+1)$-th iteration are determined according to the following rules.
 
@@
 \item 
-Перенумеровать точки $x^1,...,x^k$ предшествующих испытаний нижними индексами в порядке увеличения значений координаты
+Renumber the points $x^1,...,x^k$ of the previous trials with subscripts in ascending order of coordinate values
 \begin{equation}\label{Eq:17}
@@
 \end{equation}
-и сопоставить им значения $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, из (\ref{trial_result}), вычисленные в этих точках; точки $x_0=0$ and $x_{k+1}=1$ введены дополнительно; значения $z_0$ и $z_{k+1}$ не заданы.
+and assign $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$; points $x_0=0$ and $x_{k+1}=1$ are introduced additionally, values $z_0$ и $z_{k+1}$ are not set.
 \item
-Провести классификацию номеров  $i,1\leq i \leq k$, точек испытаний из  (\ref{Eq:17}) по числу ограничений задачи, выполняющихся в этих точках, путем построения множеств
+Classify indexes  $i,1\leq i \leq k$, of the trial points from  (\ref{Eq:17}) by the number of constraints that are met at these points by constructing the sets
 \begin{equation}\label{Eq:18}
@@
 \end{equation}
-содержащих номера всех точек  $x_i,1\leq i \leq k$, имеющих индексы, равные одному и тому же значению $\nu$.
-Граничные точки $x_0=0$ and $x_{k+1}=1$ интерпретируются как имеющие нулевые индексы, и им сопоставляется дополнительное множество $I_0={0,k+1}$.
+containing the indices of all points $x_i,1\leq i \leq k$, with indices equal to the same value $\nu$.
+The boundary points $x_0=0$ and $x_{k+1}=1$ are interpreted as having zero indices, and an additional set $I_0={0,k+1}$ is assigned to them.
 
-Определить максимальное значение индекса
+Determine the maximum value of indices
 \begin{equation}\label{Eq:19}
@@
 \end{equation}
-Если множество  $I_\nu$ содержит менее двух элементов или если $\mu_\nu$ from (\ref{Eq:20}) оказывается равным нулю, то использовать значение $\mu_\nu=1$.
+If set $I_\nu$ contains less than two elements or if $\mu_\nu$ from (\ref{Eq:20}) is equal to zero, then use the value $\mu_\nu=1$.
 \item
-Для всех непустых множеств  $I_\nu$, $1 \leq \nu \leq m+1$, вычислить значения
+For all non-empty sets $I_\nu$, $1 \leq \nu \leq m+1$, calculate values
 \begin{equation}\label{Eq:21}
@@
 \end{equation}
-где $V$ является максимальным значением индекса, а вектор
+where $V$ is the maximum index value and the vector
 \begin{equation}\label{Eq:22}
@@
 \end{equation}
-с положительными компонентами (называемый также \textit{reserve vector}) является параметром алгоритма.
+with positive components (also referred to as \textit{reserve vector}) is the algorithm parameter.
 \item
-Для каждого интервала  $(x_{i-1},x_i)$,$1 \leq i \leq k+1$, вычислить \textit{характеристику} $R(i)$: 
+For each interval $(x_{i-1},x_i)$,$1 \leq i \leq k+1$, calculate the \textit{characteristics} $R(i)$: 
 \begin{eqnarray}\label{R}
@@
 &R(i)= 2\Delta_i-4\frac{z_i-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu(x_{i-1})<\nu(x_i)=\nu,\\
-&R(i)= 2\Delta_i-4\frac{z_{i-1}-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu = \nu(x_{i-1})>\nu(x_i).\nonumber
+&R(i)= 2\Delta_i-4\frac{z_{i-1}-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu = \nu(x_{i-1})>\nu(x_i),\nonumber
 \end{eqnarray}
-где $\Delta_i=(x_i-x_{i-1})^{1/N}$, а значения $r_\nu>1, 1\leq\nu\leq m+1$, являются параметрами алгоритма.
+where $\Delta_i=(x_i-x_{i-1})^{1/N}$, and values $r_\nu>1, 1\leq\nu\leq m+1$, are the algorithm parameters.
 \item
-Упорядочить характеристики $R(i)$, $1\leq i \leq k+1$, по убыванию  	
+Sort characteristics $R(i)$, $1\leq i \leq k+1$, in descending order 	
 \begin{equation}\label{Eq:23}
@@
 \end{equation}
-и выбрать $p$ интервалов с номерами $t_j, 1\leq j \leq p$, которым соответствуют наибольшие характеристики.
+and select $p$ intervals with numbers $t_j, 1\leq j \leq p,$ corresponding to the largest characteristics.
 \item
-Провести параллельно $p$ новых испытаний в точках $x^{k+j}, 1 \leq j \leq p$, вычисленных по формулам
+Conduct $p$ new trials in parallel at points $x^{k+j}, 1 \leq j \leq p$, calculated by the formulas
 \begin{eqnarray*}
@@
 
-Алгоритм прекращает работу, если по крайней мере для одного из номеров $t_j, 1\leq j \leq p$ выполняется условие $\Delta_{t_j}\leq \epsilon$; здесь $\epsilon>0$ есть заданная точность решения задачи.
+The algorithm stops if the condition $\Delta_{t_j}\leq \epsilon$ is satisfied at least for one of the numbers $t_j, 1\leq j \leq p$; here $\epsilon>0$ is the target accuracy of the solution.
 
-Данный способ организации параллельных вычислений имеет следующее обоснование.
-Используемые в вычислительных правилах алгоритма характеристики интервалов (\ref{R}) могут рассматриваться как некоторые оценки вероятности обнаружения в данных интервалах точки глобального минимума.
-Неравенства (\ref{Eq:23}) упорядочивают интервалы по их характеристикам, и испытания проводятся параллельно в первых $p$ интервалах, имеющих наибольшие вероятности.
+This way of organizing parallel computing has the following rationale.
+The characteristics of intervals (\ref{R}) used in the computational rules of the algorithm can be considered as some estimates of the probability of detecting a global minimum point in these intervals.
+Inequalities (\ref{Eq:23}) order the intervals according to their characteristics, and trials are carried out in parallel in the first $p$ intervals with the highest probabilities.
 
 A detailed description of the algorithm convergence theory is presented in \cite{Strongin2000,Strongin2013}.
-Роль параметров алгоритма ($r_\nu>1$ and $\epsilon_\nu>0, 1\leq\nu\leq m+1$) обсуждается в \cite{Strongin2020}.
-Изложенный алгоритм может быть также использован для решения задач многокритериальной оптимизации \cite{Gergel2020}.
+The role of the algorithm parameters ($r_\nu>1$ and $\epsilon_\nu>0, 1\leq\nu\leq m+1$) is discussed in \cite{Strongin2020}.
+The described algorithm can also be used to solve multicriteria optimization problems \cite{Gergel2020}.
 
@@
 
-Одно из важных свойств рассматривамых задач заключается в том, что во многих случаях поведение функций в задаче многоэкстремальной оптимизации является неоднородным в разных подобластях поиска.
-В некоторых подобластях значения функций могут изменяться достаточно быстро (что будет соответствовать большим значениям константы Липшица в таких подобластях), в других подобластях значения могут изменяться более плавно.
-Как результат, построение единых оценок характеристик функции для всей области поиска может не полностью соответствовать ее поведению в конкретной подобласти.
+One of the important properties of the problems under consideration is that, in many cases, the behavior of functions in the multiextremal optimization problem is inhomogeneous in different search subdomains.
+In some subdomains, the values of functions can change quite quickly, which will correspond to large values of the Lipschitz constant in such subdomains.
+In other subdomains the values can change more smoothly.
+As a result, the uniform estimates of the characteristics of a function constructed for the entire search domain may not fully correspond to its behavior in a specific subdomain.
 
-Детально проработанным направлением здесь является использование разных оценок константы Липшица для каждого поискового интервала в отдельности \cite{Sergeyev2003,Sergeyev2007}.
-Определенная сложность здесь состоит в том, что при построении локальных оценок константы Липшица неявно предполагается квадратичное поведение функции в малой окрестности исследуемой точки, что может негативно сказаться на работе метода вне этой окрестности.
-Одновременно с этим при вычислении локальных (интервальных) оценок должна учитываться и общая (глобальная) оценка константы Липшица, что приводит к использованию сложных компромиссных схем оценивания константы \cite{Sergeyev2020}.
+In this case, there is a comprehensive approach, which consists in using different estimates of the Lipschitz constant for each search interval separately \cite{Sergeyev2003,Sergeyev2007}. 
+A certain difficulty here is that the construction of the local estimates for the Lipschitz constant implies that the function behaves quadratically in a small neighborhood of the point under study, which can negatively affect the operation of the method outside this neighborhood.
+At the same time, when calculating local (interval) estimates, the general (global) estimate of the Lipschitz constant should also be taken into account, which leads to the use of complex compromise schemes for estimating the constant \cite{Sergeyev2020}.
+There is a more convenient approach that takes into account the inhomogeneity of the behavior of the problem functions in different subdomains of the search domain -- adaptive prediction of the optimum position based on the accumulated search information under the assumption that the multiextremality of the problem is weak.
+The foundations of the approach are presented in \cite{Strongin2000}, some results of solving problems on parallel computing systems are given in  \cite{Barkalov2010}.
 
-Более удобным подходом, учитывающим неоднородность поведения функций задачи в разных подобластях области поиска,  является адаптивное прогнозирование положения оптимума на основе накопленной поисковой информации в предположении слабой многоэкстремальности задачи.
-Основы подхода изложены в \cite{Strongin2000}, некоторые результаты решения задач на параллельных вычислительных системах приведены в \cite{Barkalov2010}.
+In accordance with this approach, it is possible to construct adaptive forecasting rules based on the assumption that the location of the global optimum is more likely to lie in the vicinity of the points from the series $x_i, 1 \leq i \leq k,$ which correspond to small values of the normalized differences $\omega_i=(z_i-z_\nu^*)/\mu_\nu$, where $\mu_\nu$ is from (\ref{Eq:20}), and $z_\nu^*$ is from  (\ref{Eq:21}).
 
-В соответствии с данным подходом можно построить адаптивные правила прогнозирования, основанные на предположении о том, что местонахождение глобального оптимума более вероятно в окрестностях тех точек из ряда $x_i, 1 \leq i \leq k,$ которым соответствуют малые значения нормированных разностей $\omega_i=(z_i-z_\nu^*)/\mu_\nu$, где $\mu_\nu$ из (\ref{Eq:20}), а $z_\nu^*$ из (\ref{Eq:21}).
+This hypothesis is plausible either in the case when the number of performed iterations is large enough (since the algorithm, when the convergence conditions are satisfied, generates a sequence of trials that is dense only at the points of the global minimum), or in the case when the problem is unimodal (i.e.
+when the search in the vicinity of the best current approximation allows for improvement, and such successive improvements lead to a solution).
+The first case corresponds to the final phase of the search, and then the use of adaptive forecast can be interpreted as a way to speed up the refinement of the solution.
+While, the second case actually includes assumptions about the weak multiextremity of the problem.
 
-Принятая гипотеза правдоподобна либо в случае, когда число осуществленных итераций достаточно велико (т.к.
-алгоритм при выполнении условий сходимости порождает последовательность испытаний, плотную лишь в точках глобального минимума), либо в случае, когда задача является унимодальной (т.е.
-когда поиск в окрестности лучшего текущего приближения позволяет достичь улучшения, и такие последовательные улучшения ведут к решению).
-Первый случай соответствует заключительной фазе поиска, и тогда использование адаптивного прогноза можно интерпретировать как способ ускорения уточнения решения.
-Второй случай фактически включает в себя предположения о слабой многоэкстремальности задачи.
-
-Новые формулы для вычисления характеристик поисковых интервалов $R'(i)$, которые должны быть использованы на 5-м шаге алгоритма, могут быть представлены в виде 
+New formulas for calculating the characteristics of search intervals $R'(i)$, which should be used at the 5th step of the algorithm, can be written as
 \[
@@
 \]
-где $R(i)$ из (\ref{R}), а множитель $h_i$ вычисляется по правилам
+where $R(i)$ from (\ref{R}), and a coefficient $h_i$ is calculated  according to the rule
 \begin{eqnarray}\label{lR}
-& h_i = \left[\frac{\sqrt{(z_i-z_\nu^*)(z_{i-1}-z_\nu^*)}}{\mu_\nu}+1.5^{-q}\right]^{-1}, \; \nu(x_{i-1})=\nu(x_{i}),\nonumber\\
-& h_i = \left[\frac{z_i-z_\nu^*}{\mu_\nu}+1.5^{-q}\right]^{-1}, \; \nu(x_{i-1})<\nu(x_{i}),\\
-& h_i = \left[\frac{z_{i-1}-z_\nu^*}{\mu_\nu}+1.5^{-q}\right]^{-1}, \; \nu(x_{i-1})>\nu(x_{i}).\nonumber
+& h_i = \left[\frac{\sqrt{(z_i-z_\nu^*)(z_{i-1}-z_\nu^*)}}{\mu_\nu}+10^{-q}\right]^{-1}, \; \nu(x_{i-1})=\nu(x_{i}),\nonumber\\
+& h_i = \left[\frac{z_i-z_\nu^*}{\mu_\nu}+10^{-q}\right]^{-1}, \; \nu(x_{i-1})<\nu(x_{i}),\\
+& h_i = \left[\frac{z_{i-1}-z_\nu^*}{\mu_\nu}+10^{-q}\right]^{-1}, \; \nu(x_{i-1})>\nu(x_{i}).\nonumber
 \end{eqnarray}
-Здесь целочисленный неотрицательный параметр $q$ характеризует «степень локальности» поиска: при $q=0$ поиск будет иметь ``глобальный'', а при $q>>1$ -- ``локальный'' характер. 
-В самом деле, множитель $h_i$ будет принимать наибольшее значение равное $1.5^q$ лишь для тех интервалов, граничные точки которых соответствуют текущему минимуму, т.е.
-для которых выполняется равенство $\omega_i = 0$.
-Указанное значение множителя $h_i$ будет большим при больших значениях параметра $q$ (что увеличит значение характеристики данного интервала), и будет равно 1 при $q=0$ (что соответствует исходному значению характеристики).
+Here, the non-negative integer parameter $q$ characterizes the ``degree of locality'' of the search: for $q=0$ the search will have a ``global'' nature, and for $q>1$ -- ``local''. 
+Indeed, the coefficient $h_i$ will have the largest value equal to $10^q$ only for those intervals the boundary points of which correspond to the current minimum, i.e.
+for which the equality $\omega_i = 0$ holds.
+The specified value of the coefficient $h_i$ will be higher for larger values of the parameter $q$ (which will increase the value of the characteristic of this interval), and will be equal to 1 for $q=0$ (which corresponds to the initial value of the characteristics).
 
-Управление параметром $q$, входящим в (\ref{lR}), позволяет учитывать различные предположения о характере экстремума.
-Например, одной из стратегий управления является чередование с заданной частотой больших и малых значений $q$, что соответствует смеси локального и глобального поиска, которую можно интерпретировать как сочетание процедуры локального уточнения текущего решения и процедуры глобального обновления этого решения.
-При этом в течение первых итераций рекомендуется использовать значение $q=0$, т.к.
-накопленная на них поисковая информация обеспечит адекватность формул (\ref{lR}) при последующем локальном уточнении.
+Controlling the parameter $q$, included in (\ref{lR}), allows taking into account various assumptions about the nature of the extremum.
+For example, one of the control strategies is the alternation with a given frequency of large and small values of $q$, corresponding to a mixture of local and global search, which can be interpreted as a combination of local refinement of the current solution and global updating of this solution.
+In this case, during the first iterations, it is recommended to use the value $q=0$, since the search information accumulated during these iterations will ensure the adequacy of formulas (\ref{lR}) after the subsequent local refinement.
 
@@
 
-Рассмотрим теперь случай, когда аргумент функций задачи содержит две компоненты: вектор $y$, принадлежащий гиперинтервалу $D$, и вектор $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е.
+Let us now consider the case when the argument of the problem functions contains two components: vector $y$, which belongs to the hyperinterval $D$, and vector $u$, which has a finite (and not very large) set of possible values, i.e.
 \begin{eqnarray}\label{problem_i}
@@
 
-Такие конечные наборы могут характеризовать, например, вариант материала, из которого создается объект, геометрические размеры или другие величины, которые могут принадлежать стандартному дискретному множеству, и т.п.
+Such finite sets can characterize, for example, a variation of the material from which an object is created, geometric dimensions or other quantities that can belong to a standard discrete set, etc.
 
-Занумеруем целыми числами $s, 1\leq s \leq S,$ все возможные значения вектора $u$, т.е. сопоставим каждому рассматриваемому значению $s$ вектор $u_s$. 
-Тогда рассматриваемая задача может быть записана в виде 
+Let us assign integers  $s, 1\leq s \leq S,$ to all possible values of the vector $u$, i.e. associate each considered value $s$ with a vector $u_s$. 
+Then the problem can be written as
 \begin{eqnarray}\label{problem_is}
@@
 
-Используя схему редукции размерности с помощью Hilbert curve $y(x), x\in [0,1]$ можно сопоставить каждой задаче минимизации по $y$ одномерную задачу минимизации
+Using the dimensionality reduction scheme by the Hilbert curve $y(x), x\in [0,1]$ we can associate each minimization problem in $y$ with a one-dimensional minimization problem
 \[
@@
 
-Рассмотрим теперь отображение 
+Now let us consider the mapping
 \[
@@
 \]
-ставящее в соответствие точке из интервала [0,S] точку в области $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функции 
+Which assigns a point on the interval  [0,S] to a point in the domain $D$ ($E(x)$ corresponds to the integer part of the number $x$) and define the functions 
 \[
@@
 \]
-имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_k = i, 1\leq i \leq S-1$.
-Значение  $z_k = g_\nu(y(x_k))$ из (\ref{trial_result}) в этих точках будем считать неопределенным, а значения индекса -- равным 0, т.е. $\nu(x_k) = 0$ .
+which have discontinuities of the first kind at integer points $x_k = i, 1\leq i \leq S-1$.
+The value  $z_k = g_\nu(y(x_k))$ from (\ref{trial_result}) at these points will be considered undefined, and the values of the index -- equal to 0, i.e.  $\nu(x_k) = 0$ .
 
-Используя введенные обозначения можно переформулировать исходную задачу как
+Using the above assignments, we can reformulate the original problem as
 \begin{equation}\label{problem_is1}
@@
 
-Для иллюстрации на Fig. \ref{fig:1}  изображены графики функций, соответствующих задаче с одним непрерывным и одним бинарным параметром.
+To illustrate this, Fig. \ref{fig:1} shows graphs of functions corresponding to a problem with one continuous and one binary parameter.
 \begin{figure}[ht]
@@
 
-Применяя к решению задачи (\ref{problem_is1}) параллельный индексный алгоритм с локальной настройкой, мы найдем решение задачи (\ref{problem_i}).
-При этом основная часть испытаний будет проведена в той подзадаче, решение которой соответствует решению исходной задачи (\ref{problem_i}).
-В остальных подзадачах будет проведена лишь незначительная часть испытаний, т.к.
-решения данных подзадач являются локально-оптимальными по отношению к решению $s$-й подзадачи.
-Сказанное подтверждается Fig.
-\ref{fig:1}, где штрихами обозначены точки испытаний, выполненных при решении данной задачи.
+By applying a parallel index algorithm with local tuning to the solution of problem  (\ref{problem_is1}), ), we find a solution to the problem  (\ref{problem_i}).
+In this case, the main part of the trials will be carried out in the subproblem, the solution of which corresponds to the solution of the original problem (\ref{problem_i}).
+).
+In the remaining subtasks, only an insignificant part of the trials will be carried out, since solutions of these subproblems are locally optimal for the solution of the xth subproblem.
+This is confirmed by Fig.
+\ref{fig:1}, where the dashed lines indicate the trial points performed during the solution of this problem.
 
-Таким образом, мы сформировали Parallel Mixed-integer Global search Algorithm with Local tuning (PMGAL), основанный на сведении задачи mixed-integer non-convex optimization problem к одновременно решаемому набору непрерывных задач.
+Thus, we have formed the Parallel Mixed-integer Global search Algorithm with Local tuning (PMGAL), based on the reduction of the mixed-integer non-convex optimization problem to a simultaneously solved set of continuous problems.
 
-Предложенная схема решения mixed-integer global optimization problems основана на параллельном индексном методе для решения задач непрерывной оптимизации и не является ориентированной на конкретное вычислительное устройство.
-Одной из ключевых операций здесь является параллельное проведение нескольких испытаний в различных точках области поиска (см.
-шаг 7 алгоритма), которое может быть реализовано как на CPU (с использованием OpenMP и/или MPI), так и на GPU (с использованием CUDA).
-В данной работе для распараллеливания были задействованы только центральные процессоры.
-Распределение работ между узлами осуществлялось в соответствии с многоуровневой схемой декомпозиции параллельных вычислений, изложенной в \cite{Strongin2018,Barkalov2020}.
+The proposed scheme for solving mixed-integer global optimization problems is based on a parallel index method for solving continuous optimization problems and is not oriented to a specific computing device.
+One of the key operations here is the parallel execution of several trials at different points of the search area (see step 7 of the algorithm), which can be implemented both on the CPU (using OpenMP and / or MPI) and on the GPU (using CUDA).
+In this work, parallelization was done only on CPUs.
+The workload was distributed between the nodes in accordance with the multilevel parallel computation decomposition scheme described in \cite{Strongin2018,Barkalov2020}.
 
@@
 
-%Ниже пока переводить не надо
+%Переводить до конца, т.е. до acknowledgments
 
-Первая серия экспериментов была проведена с использованием последовательной версии предложенного алгоритма с целью его сравнения с известными методами аналогичного назначения.
+The first series of experiments was carried out using a sequential version of the algorithm without using local tuning in order to compare it with other similar methods.
 
-Предложенный в данной работе алгоритм PMGAL (в его последовательном варианте) сравнивался с genetic algorithm for solving mixed-integer optimization problems, реализованным в Matlab Global Optimization Toolbox.
-В Table \ref{tab:1} приведено число испытаний, потребовавшееся для решения известных тестовых mixed-integer problems данными методами.
-Для обоих методов была использована одинаковая точность поиска решения $10^{-2}$.
-Все вычислительные эксперименты были проведены на компьютере с процессором Intel Core i5-7300 2.5 GHz и 8 Gb RAM.
-Результаты экспериментов показывают превосходство метода PMGAL как по числу итераций, так и по времени работы.
+The PMGAL algorithm proposed in this work (in its sequential version) was compared with the genetic algorithm for solving mixed-integer optimization problems implemented in the Matlab Global Optimization Toolbox.
+Table \ref{tab:1} shows the number of trials it took for these methods to solve a number of mixed-integer trial problems taken from \cite{Deep,Floudas}.
+The same solution search accuracy $10^{-2}$ was used for both methods.
+All computational experiments were carried out on a computer with an Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM.
+The experimental results show the superiority of the PMGAL method both in terms of the number of iterations and the running time.
 
 \begin{table}
-	\caption{Сравнение эффективности методов PMGAL и GA}
+	\caption{Comparison of the effectiveness of the PMGAL and GA methods}
 	\label{tab:1}
@@
 
-%Последующие серии экспериментов были выполнены для оценки ускорения параллельной версии предложенного алгоритма MIPIA с использованием как CPU, так и GPU.
-В данных экспериментах решалась серия из 100 тестовых mixed-integer problems, сгенерированных случайных образом.
-
-Известным генератором тестовых задач непрерывной многоэкстремальной оптимизации является GKLS-генератор \cite{Gaviano}.
-Он позволяет генерировать тестовые функции заданной размерности с заданными свойствами: известным числом локальных минимумов, известной областью притяжения глобального решения, и т.д.
+The next experiment was conducted to assess the effect of local tuning on accelerating the convergence of the algorithm.
+We used the GKLS generator \cite{Gaviano} to generate trial problems of continuous multiextremal optimization.
+GKLS allows generating continuous trial functions of a given dimensionality with specific properties: a known number of local minima, a known domain of attraction of the global solution, etc.
 %This generator of multiextremal functions is often used for the investigations of the global optimization algorithms \cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
 
-In Fig.~\ref{example} (a) and (b), the contour plots of two-dimensional GKLS functions are presented.
-Figures also shows the points of the trials performed by the method until the required accuracy $\epsilon=10^{-2}$ was achieved.
+Fig.~\ref{example} (a) and (b) shows the contour plots of two-dimensional GKLS functions.
+These figures also shows the points of the trials performed by the global search until the required accuracy $\epsilon=10^{-2}$ was achieved.
 
@@
 
-В проведенных экспериментах GKLS-генератор использовался как основа для построения mixed-integer problems. Правила, позволяющие генерировать тестовые задачи подобного вида, заключаются в следующем.
+In our experiments, the GKLS generator was used as a basis for constructing mixed-integer problems. The rules for generating trial problems of this type were as follows.
 
 \begin{enumerate}
-	\item С помощью генератора GKLS порождается непрерывная многоэкстремальная функция $\psi(y), \; y\in D, \; D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\}$.
-Глобальный минимум данной функции достигается в известной точке $y'=(y'_1,...,y'_N)$ и равен $\psi'=\psi(y')=-1$.
-	\item Генерируется convex mixed-integer function 
+	\item GKLS generates a continuous multiextremal function $\psi(y), \; y\in D, \; D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\}$.
+The global minimum of this function is achieved at the known point $y'=(y'_1,...,y'_N)$ and is equal to $\psi'=\psi(y')=-1$.
+	\item GKLS generates a convex mixed-integer function
 	\[
@@
 	\]
-	где 
+	where 
 	\begin{eqnarray*}
@@
 	\end{eqnarray*}
-	т.е. данная функция имеет $N$ непрерывных и $M$ дискретных параметров и достигает минимального значения в точке $(y',b)$.
-	\item Вычисляется коэффициент 
+	i.e. this function has $N$ continuous and $M$ discrete parameters and reaches its minimum value at the point $(y',b)$.
+	\item The coefficient is calculated
 	\[
-	C = 4 - \min_{y,u} \left\{ \phi(y,u) \right\}, \; y\in D, u \in U.
+	C = 4 - \max_{y,u} \left\{ \phi(y,u) \right\}, \; y\in D, u \in U.
 	\]
-	Минимум вогнутой функции $-\phi(y,u)$ достигается в одной из угловых точек области поиска, поэтому при размерности задачи порядка 10 его вычисление может быть выполнено методом перебора.
-	\item Формируется многоэкстремальная mixed-integer function 
+	The maximum of the convex function $\phi(y,u)$ is achieved at one of the corner points of the search domain; therefore, when the dimensionality of the problem is around 10, its calculation can be performed by the enumeration method.
+	\item  multi-extreme mixed-integer function is formed
 	\[
-	\varphi(y,u) = \left(\psi(y) + \sum_{j=1}^M{u_j}\right)\left(C - \phi(y,u)\right).
+	\varphi(y,u) = \left(\psi(y) - \sum_{j=1}^M{u_j}\right)\left(C - \phi(y,u)\right).
 	\]
-	По построению $\varphi(y,u)$  будет принимать минимальное значение в точке $(y',b)$.
+	After mapping $\varphi(y,u)$  will reach the minimum value at the point  $(y',b)$.
 
@@
 
-In the problems generated in our experiments, there were 5 discrete and 6 continuous parameters 
+The problems generated in our experiments had 5 discrete and 4 continuous parameters 
 	\begin{eqnarray*}
-	& y\in D = \left\{ -1 \leq y_j\leq 1, 1\leq j \leq 6 \right\} \subset R^6, \\
+	& y\in D = \left\{ -1 \leq y_j\leq 1, 1\leq j \leq 4 \right\} \subset R^4, \\
 	& u\in U = \left\{ u_j \in  \left\{-1, -1/3, 0, 1/3, 1 \right\}, 1\leq j \leq 5 \right\}.
@@
 
+In total, we generated and solved 100 9-dimensional mixed-integer problems of this type.
+In order to simulate the computational complexity inherent in applied optimization problems, the calculation of the function in all the experiments was complicated by additional calculations that did not change the form of the function and the location of its minima.
+The accuracy of the search was equal to $10^{-2}$.
+Computational experiments were carried out on the Lobachevsky supercomputer.
+The node of supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.
+
+Table \ref{tab:2} summarizes the experimental results obtained for the sequential version of the PMGAL algorithm using different strategies for alternating local and global search.
+For local search in formulas  (\ref{lR}), we used the parameter $q=3$, for global -- $q=0$.
+The first column of the table shows the frequency of alternation of the indicated values.
+For example, 1:1 corresponds to alternating local and global search at each iteration; 1:2 corresponds to one local search step and two global search steps, and so on.
+The second column of the table reflects the average number of iterations  $K_{av}$ (thousand.), required to solve a problem with a given accuracy.
+
+\begin{table}
+	\caption{Comparison of local tuning strategies}
+	\label{tab:2}
+	\center
+	\begin{tabular}{|c|c|}
+		\hline		
+		\textit{loc} : \textit{glob} & $K_{av}$   \\
+		\hline 
+		1 : 1 & 102.6\\
+		1 : 2 & 89.2\\
+		1 : 4 & 97.5\\
+		1 : 8 & 127.1\\
+		1 : 16 & 144.2\\
+		\hline
+	\end{tabular}
+\end{table}
+
+Experimental results show that the least number of iterations is required for the locally tuned method when using the 1:2 alternation strategy, i.e.
+one local refinement step alternates with two global search steps.
+We will use this strategy in further experiments.
+
+\begin{table}
+	\caption{Average running time $T_{av}$ of the parallel algorithm with a local setting}
+	\label{tab:3}
+	\center
+	\begin{tabular}{|c|c|c|c|}
+		\hline	
+	\diaghead{\theadfont Diag Column} { $N_{p}$}{$N_{t}$} & 1 & 4 & 8\\
+
+		\hline		
+		2  & 199.0 & 65.4 & 61.0  \\
+		5  & 91.2 & 28.0 &  26.1 \\
+		17  & 37.2 & 10.8 & 7.5 \\
+		65  & 29.3 & 8.7 & 5.4 \\
+		\hline
+	\end{tabular}
+\end{table}
+
+The last series of experiments was performed to evaluate the acceleration of the parallel version of the PMGAL algorithm. 
+Table \ref{tab:3} shows the average solution time  $T_{av}$ (in seconds) depending on the involved number of processes $N_{p}$ and threads $N_{t}$ in each process.
+Our experiments involved from 1 to 16 cluster nodes depending on the number of parallel processes.
+The acceleration provided by the parallel algorithm is not ideal, but quite acceptable.
+
+\section{Conclusions}
+
+This article presents the results related to the development of parallel algorithms for global optimization for solving multiextremal problems in which some of the parameters are continuous, and some are discrete.
+These problems can also include non-convex constraints, which are accounted for using a special index scheme.
+We propose to solve the problems of this class using a parallel algorithm.
+The basic principle of organizing parallel computations in the algorithm is based on the simultaneous carrying out of several trials at different points of the search domain.
+
+The parallel global optimization algorithm uses a local tuning technique based on the assumption that the multiextremity of the problem is weak.
+This assumption is true, in particular, at the final phase of the search, when one or more trials have already been performed in the vicinity of the global solution, and the only remaining step is to refine the found estimate of the solution.
+
+In its sequential form, the proposed algorithm remains on par with the similar methods implemented in Matlab Global Optimization Toolbox. 
+In a parallel form, the proposed algorithm demonstrates acceptable scalability up to dozens of parallel processes.
+
 \begin{acknowledgments}
-%We are grateful to C.~C.~Surname3 and the reviewers for careful reading of the manuscript and helpful remarks.
 This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.

### 2021-02-18 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
Add Gergel as coauthor
Поправил формулу и заменил русскую букву

@@
 %\authorrunning{First-Author at al.} % for running heads
-\authorrunning{Barkalov, Lebedev} % for running heads
+\authorrunning{Barkalov, Gergel, Lebedev} % for running heads
 
@@
 \email[E-mail: ]{konstantin.barkalov@itmm.unn.ru}
+\affiliation{Lobachevsky State University of Nizhni Novgorod, Gagarin ave. 23, Nizhni Novgorod, 603950 Russia}
+
+\author{\firstname{V.~P.}~\surname{Gergel}}
+\email[E-mail: ]{gergel@unn.ru}
 \affiliation{Lobachevsky State University of Nizhni Novgorod, Gagarin ave. 23, Nizhni Novgorod, 603950 Russia}
@@
 Section 3 outlines a way to take into account the local properties of the problem functions to speed up the convergence of the algorithm.
-
-В Section 4 предложен подход к обобщению параллельного индексного алгоритма с целью решения mixed-integer problems.
-Приведен способ, позволяющий свести решение mixed-integer problem к решению информационно-связанного множества задач непрерывной оптимизации, которое может выполняться параллельно.
-
+Section 4 proposes an approach to generalize the parallel index algorithm to solve mixed-integer problems.
+This method allows reducing the solution of the mixed-integer problem to the solution of an information-related set of continuous optimization problems which can be performed in parallel.
 Section 5 contains the results of numerical experiments.
@@
 
-\section{Parallel global optimization algorithm and dimensionality reduction}
+\section{Parallel global optimization algorithm\protect\\
+and dimensionality reduction}
 
@@
 \end{equation}
-and assign $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$; points $x_0=0$ and $x_{k+1}=1$ are introduced additionally, values $z_0$ и $z_{k+1}$ are not set.
+and assign $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$; points $x_0=0$ and $x_{k+1}=1$ are introduced additionally, values $z_0$ and $z_{k+1}$ are not set.
 \item
@@
 
-Using the dimensionality reduction scheme by the Hilbert curve $y(x), x\in [0,1]$ we can associate each minimization problem in $y$ with a one-dimensional minimization problem
+Using the dimensionality reduction scheme by the Peano-Hilbert curve $y(x), x\in [0,1],$ we can associate each minimization problem in $y$ with a one-dimensional minimization problem
 \[
@@
 \]
-Which assigns a point on the interval  [0,S] to a point in the domain $D$ ($E(x)$ corresponds to the integer part of the number $x$) and define the functions 
+which assigns a point on the interval  [0,S] to a point in the domain $D$ ($E(x)$ corresponds to the integer part of the number $x$) and define the functions 
 \[
@@
 \]
-which have discontinuities of the first kind at integer points $x_k = i, 1\leq i \leq S-1$.
-The value  $z_k = g_\nu(y(x_k))$ from (\ref{trial_result}) at these points will be considered undefined, and the values of the index -- equal to 0, i.e.  $\nu(x_k) = 0$ .
+which have jump discontinuities at integer points $x_k = i, 1\leq i \leq S-1$.
+The values $z_k = g_\nu(y(x_k))$ from (\ref{trial_result}) at these points will be considered undefined, and the values of the index -- equal to 0, i.e. $\nu(x_k) = 0$ .
 
@@
 
-By applying a parallel index algorithm with local tuning to the solution of problem  (\ref{problem_is1}), ), we find a solution to the problem  (\ref{problem_i}).
+By applying a parallel index algorithm with local tuning to the solution of problem  (\ref{problem_is1}), we find a solution to the problem  (\ref{problem_i}).
 In this case, the main part of the trials will be carried out in the subproblem, the solution of which corresponds to the solution of the original problem (\ref{problem_i}).
-).
-In the remaining subtasks, only an insignificant part of the trials will be carried out, since solutions of these subproblems are locally optimal for the solution of the xth subproblem.
+In the remaining subtasks, only an insignificant part of the trials will be carried out, since solutions of these subproblems are locally optimal.
 This is confirmed by Fig.
@@
 The proposed scheme for solving mixed-integer global optimization problems is based on a parallel index method for solving continuous optimization problems and is not oriented to a specific computing device.
-One of the key operations here is the parallel execution of several trials at different points of the search area (see step 7 of the algorithm), which can be implemented both on the CPU (using OpenMP and / or MPI) and on the GPU (using CUDA).
+One of the key operations here is the parallel execution of several trials at different points of the search domain (see step 7 of the algorithm), which can be implemented both on the CPU (using OpenMP and / or MPI) and on the GPU (using CUDA).
 In this work, parallelization was done only on CPUs.
-The workload was distributed between the nodes in accordance with the multilevel parallel computation decomposition scheme described in \cite{Strongin2018,Barkalov2020}.
+The workload was distributed between the nodes in accordance with the multilevel scheme of parallel computation decomposition described in \cite{Strongin2018,Barkalov2020}.
 
 \section{Results of experiments}
-
-%Переводить до конца, т.е. до acknowledgments
 
@@
 The PMGAL algorithm proposed in this work (in its sequential version) was compared with the genetic algorithm for solving mixed-integer optimization problems implemented in the Matlab Global Optimization Toolbox.
-Table \ref{tab:1} shows the number of trials it took for these methods to solve a number of mixed-integer trial problems taken from \cite{Deep,Floudas}.
-The same solution search accuracy $10^{-2}$ was used for both methods.
+Table \ref{tab:1} shows the number of trials it took for these methods to solve a number of mixed-integer test problems taken from \cite{Deep,Floudas}.
+The same accuracy of search $10^{-2}$ was used for both methods.
 All computational experiments were carried out on a computer with an Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM.
@@
 The next experiment was conducted to assess the effect of local tuning on accelerating the convergence of the algorithm.
-We used the GKLS generator \cite{Gaviano} to generate trial problems of continuous multiextremal optimization.
-GKLS allows generating continuous trial functions of a given dimensionality with specific properties: a known number of local minima, a known domain of attraction of the global solution, etc.
+We used the GKLS generator \cite{Gaviano} to generate test problems of continuous multiextremal optimization.
+GKLS allows generating continuous functions of a given dimensionality with specific properties: a known number of local minima, a known domain of attraction of the global solution, etc.
 %This generator of multiextremal functions is often used for the investigations of the global optimization algorithms \cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
@@
 The global minimum of this function is achieved at the known point $y'=(y'_1,...,y'_N)$ and is equal to $\psi'=\psi(y')=-1$.
-	\item GKLS generates a convex mixed-integer function
+	\item A convex mixed-integer function
 	\[
-			\phi (y,u) = 2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} \right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right],
+			\phi (y,u) = 2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} \right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right]
 	\]
-	where 
+	is generated, where 
 	\begin{eqnarray*}
@@
 	\]
-	The maximum of the convex function $\phi(y,u)$ is achieved at one of the corner points of the search domain; therefore, when the dimensionality of the problem is around 10, its calculation can be performed by the enumeration method.
-	\item  multi-extreme mixed-integer function is formed
+	The maximum of the convex function $\phi(y,u)$ is achieved at one of the boundary points of the search domain; therefore, when the dimensionality of the problem is around 10, its calculation can be performed by the enumeration method.
+	\item  Multi-extreme mixed-integer function is formed
 	\[
@@
 	\]
-	After mapping $\varphi(y,u)$  will reach the minimum value at the point  $(y',b)$.
+	By construction, $\varphi(y,u)$ will reach the minimum value at the point $(y',b)$.
 
@@
 	& y\in D = \left\{ -1 \leq y_j\leq 1, 1\leq j \leq 4 \right\} \subset R^4, \\
-	& u\in U = \left\{ u_j \in  \left\{-1, -1/3, 0, 1/3, 1 \right\}, 1\leq j \leq 5 \right\}.
+	& u\in U = \left\{ u_j \in  \left\{-1, -1/3, 1/3, 1 \right\}, 1\leq j \leq 4 \right\}.
 	\end{eqnarray*}
@@
 Table \ref{tab:2} summarizes the experimental results obtained for the sequential version of the PMGAL algorithm using different strategies for alternating local and global search.
-For local search in formulas  (\ref{lR}), we used the parameter $q=3$, for global -- $q=0$.
+For local search in formulas  (\ref{lR}), we used the parameter $q=3$, for global search -- $q=0$.
 The first column of the table shows the frequency of alternation of the indicated values.
@@
 
-Experimental results show that the least number of iterations is required for the locally tuned method when using the 1:2 alternation strategy, i.e.
+Experimental results show that the least number of iterations is required for the method when using the 1:2 alternation strategy, i.e.
 one local refinement step alternates with two global search steps.
@@
 \begin{table}
-	\caption{Average running time $T_{av}$ of the parallel algorithm with a local setting}
+	\caption{Average running time $T_{av}$ of the parallel algorithm with a local tuning}
 	\label{tab:3}
@@
 These problems can also include non-convex constraints, which are accounted for using a special index scheme.
-We propose to solve the problems of this class using a parallel algorithm.
+We propose a parallel algorithm for solving the problems of this class.
 The basic principle of organizing parallel computations in the algorithm is based on the simultaneous carrying out of several trials at different points of the search domain.

### 2021-02-18 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
fix some misprints

@@
 
-\firstcollaboration{(Submitted by A.~A.~Editor-name)} % Add if you know submitter.
+\firstcollaboration{(Submitted by V.~V.~Voevodin)} % Add if you know submitter.
 %\lastcollaboration{ }
 
-\received{February 1, 2021} % The date of receipt to the editor, i.e. December 06, 2017
+\received{February 20, 2021} % The date of receipt to the editor, i.e. December 06, 2017
 
@@
 An extensive literature is devoted to methods for solving mixed-integer problems (see, for example, reviews \cite{Burer,Boukouvala}).
-Known deterministic methods for solving problems of this class are based, as a rule, on the Branch-and-Bound \cite{Belotti} or Branch-and-Reduce \cite{Vigerske}.
-.
+Known deterministic methods for solving problems of this class are based, as a rule, on the Branch-and-Bound \cite{Belotti} or Branch-and-Reduce approach \cite{Vigerske}.
 Such problems may also be solved by nature-inspired algorithms \cite{Deep,Schluter}, which in one way or another are based on a random search.
@@
 
-Within this approach 
+Within this approach: 
 \begin{itemize}
@@
 Let us consider a parallel algorithm for solving constrained global optimization problems proposed in  \cite{Strongin2000,Strongin2013} and based on an index scheme for the accounting of constraints.
-
 To describe a parallel algorithm, we will assume that we have $p$ computational elements (nodes, processors, or cores) at our disposal, which will be used for parallel execution of $p$ trials simultaneously.
@@
 containing the indices of all points $x_i,1\leq i \leq k$, with indices equal to the same value $\nu$.
-The boundary points $x_0=0$ and $x_{k+1}=1$ are interpreted as having zero indices, and an additional set $I_0={0,k+1}$ is assigned to them.
+The boundary points $x_0=0$ and $x_{k+1}=1$ are interpreted as having zero indices, and an additional set $I_0=\{0,k+1\}$ is assigned to them.
 
@@
 Conduct $p$ new trials in parallel at points $x^{k+j}, 1 \leq j \leq p$, calculated by the formulas
-\begin{eqnarray*}
-& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}, \; \nu(x_{t_j-1})\neq \nu(x_{t_j}), \\
-& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}- \frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \; \nu(x_{t_j-1})=\nu(x_{t_j})=\nu. \\
-\end{eqnarray*}
+\begin{gather*}
+x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}, \; \nu(x_{t_j-1})\neq \nu(x_{t_j}), \\
+x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}- \frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \; \nu(x_{t_j-1})=\nu(x_{t_j})=\nu. \\
+\end{gather*}
 
@@
 
-In accordance with this approach, it is possible to construct adaptive forecasting rules based on the assumption that the location of the global optimum is more likely to lie in the vicinity of the points from the series $x_i, 1 \leq i \leq k,$ which correspond to small values of the normalized differences $\omega_i=(z_i-z_\nu^*)/\mu_\nu$, where $\mu_\nu$ is from (\ref{Eq:20}), and $z_\nu^*$ is from  (\ref{Eq:21}).
+In accordance with this approach, it is possible to construct adaptive forecasting rules based on the assumption that the location of the global optimum is more likely to lie in the vicinity of the points from the series $x_i, 1 \leq i \leq k,$ from (\ref{Eq:17}), which correspond to small values of the normalized differences $\omega_i=(z_i-z_\nu^*)/\mu_\nu$, where $\mu_\nu$ is from (\ref{Eq:20}), and $z_\nu^*$ is from  (\ref{Eq:21}).
 
 This hypothesis is plausible either in the case when the number of performed iterations is large enough (since the algorithm, when the convergence conditions are satisfied, generates a sequence of trials that is dense only at the points of the global minimum), or in the case when the problem is unimodal (i.e.
-when the search in the vicinity of the best current approximation allows for improvement, and such successive improvements lead to a solution).
+when the search in the vicinity of the current best point allows for improvement, and such successive improvements lead to a solution).
 The first case corresponds to the final phase of the search, and then the use of adaptive forecast can be interpreted as a way to speed up the refinement of the solution.
@@
 
-Let us now consider the case when the argument of the problem functions contains two components: vector $y$, which belongs to the hyperinterval $D$, and vector $u$, which has a finite (and not very large) set of possible values, i.e.
+Let us now consider the case when the argument of the problem functions contains two components: vector $y$, which belongs to the hyperinterval $D$, and vector $u$, which has a finite (and not very large) set of possible values $U$, i.e.
 \begin{eqnarray}\label{problem_i}
-& \min{\left\{ g_{m+1}(y,u):y\in D, \; g_i(y,u)\leq 0, \; 1 \leq i \leq m\right\}},\\
+& \min{\left\{ g_{m+1}(y,u):y\in D, u \in U, \; g_i(y,u)\leq 0, \; 1 \leq i \leq m\right\}},\\
 & D=\left\{a_j\leq y_j \leq b_j, \; 1\leq j \leq N \right\}.\nonumber
@@
 Then the problem can be written as
-\begin{eqnarray}\label{problem_is}
-& \min_{s\in\{1,...,S\}}\left(\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_i(y,u_s)\leq 0, \; 1 \leq i \leq m\right\}}\right),\\
-& D=\left\{ a_j\leq y_j \leq b_j, \; 1 \leq j\leq N \right\}.\nonumber 
-\end{eqnarray}
+\begin{gather}\label{problem_is}
+ \min_{s\in\{1,...,S\}}\left\{\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_i(y,u_s)\leq 0, \; 1 \leq i \leq m\right\}}\right\},\\
+ D=\left\{ a_j\leq y_j \leq b_j, \; 1 \leq j\leq N \right\}.\nonumber 
+\end{gather}
 
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{fig1.jpg}
+    \includegraphics[width=0.9\textwidth]{fig1.jpg}
     \caption{Mixed-integer global optimization problem}
@@
 The first series of experiments was carried out using a sequential version of the algorithm without using local tuning in order to compare it with other similar methods.
-
 The PMGAL algorithm proposed in this work (in its sequential version) was compared with the genetic algorithm for solving mixed-integer optimization problems implemented in the Matlab Global Optimization Toolbox.
 Table \ref{tab:1} shows the number of trials it took for these methods to solve a number of mixed-integer test problems taken from \cite{Deep,Floudas}.
-The same accuracy of search $10^{-2}$ was used for both methods.
+The same accuracy of search $\epsilon = 10^{-2}$ was used for both methods.
 All computational experiments were carried out on a computer with an Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM.
@@
 
-\begin{table}
+\begin{table}[ht]
 	\caption{Comparison of the effectiveness of the PMGAL and GA methods}
@@
 		\hline
-	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c|}{ GA } &  \multicolumn{2}{c|}{MIGSA} \\
+	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c|}{ GA } &  \multicolumn{2}{c|}{PMGAL} \\
 		\cline{2-3} \cline{4-5} 
@@
 %This generator of multiextremal functions is often used for the investigations of the global optimization algorithms \cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}. 
-
 Fig.~\ref{example} (a) and (b) shows the contour plots of two-dimensional GKLS functions.
@@
 \hfill
-\begin{minipage}{0.48\linewidth}
+\begin{minipage}{0.475\linewidth}
 \center{\includegraphics[width=1.0\linewidth]{GKLS-6.png} \\ (b)}
@@
 	i.e. this function has $N$ continuous and $M$ discrete parameters and reaches its minimum value at the point $(y',b)$.
-	\item The coefficient is calculated
+	\item The coefficient
 	\[
-	C = 4 - \max_{y,u} \left\{ \phi(y,u) \right\}, \; y\in D, u \in U.
+	C = 4 - \max \left\{ \phi(y,u) \right\}, \; y\in D, u \in U,
 	\]
+	is calculated.
 The maximum of the convex function $\phi(y,u)$ is achieved at one of the boundary points of the search domain; therefore, when the dimensionality of the problem is around 10, its calculation can be performed by the enumeration method.
-	\item  Multi-extreme mixed-integer function is formed
+	\item  Multiextremal mixed-integer function
 	\[
-	\varphi(y,u) = \left(\psi(y) - \sum_{j=1}^M{u_j}\right)\left(C - \phi(y,u)\right).
+	\varphi(y,u) = \left(\psi(y) - \sum_{j=1}^M{u_j}\right)\left(C - \phi(y,u)\right)
 	\]
-	By construction, $\varphi(y,u)$ will reach the minimum value at the point $(y',b)$.
+	is formed. By construction, $\varphi(y,u)$ will reach the minimum value at the point $(y',b)$.
 
@@
 	& y\in D = \left\{ -1 \leq y_j\leq 1, 1\leq j \leq 4 \right\} \subset R^4, \\
-	& u\in U = \left\{ u_j \in  \left\{-1, -1/3, 1/3, 1 \right\}, 1\leq j \leq 4 \right\}.
+	& u\in U = \left\{ u_j \in  \left\{-1, -1/3, 1/3, 1 \right\}, 1\leq j \leq 5 \right\}.
 	\end{eqnarray*}
@@
 In order to simulate the computational complexity inherent in applied optimization problems, the calculation of the function in all the experiments was complicated by additional calculations that did not change the form of the function and the location of its minima.
-The accuracy of the search was equal to $10^{-2}$.
+The accuracy of the search was equal to $\epsilon = 10^{-2}$.
 Computational experiments were carried out on the Lobachevsky supercomputer.
@@
 \begin{table}
-	\caption{Average running time $T_{av}$ of the parallel algorithm with a local tuning}
+	\caption{Average running time $T_{av}$ of the parallel algorithm with local tuning}
 	\label{tab:3}

### 2021-02-18 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
Revert "Поправил формулу и заменил русскую букву"
This reverts commit 218dd72cad16e06e5218888f466113f34057398d.

--- 2021-02-18 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
@@
 \end{equation}
-and assign $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$; points $x_0=0$ and $x_{k+1}=1$ are introduced additionally, values $z_0$ and $z_{k+1}$ are not set.
+and assign $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$; points $x_0=0$ and $x_{k+1}=1$ are introduced additionally, values $z_0$ и $z_{k+1}$ are not set.
 \item
@@
 	& y\in D = \left\{ -1 \leq y_j\leq 1, 1\leq j \leq 4 \right\} \subset R^4, \\
-	& u\in U = \left\{ u_j \in  \left\{-1, -1/3, 1/3, 1 \right\}, 1\leq j \leq 4 \right\}.
+	& u\in U = \left\{ u_j \in  \left\{-1, -1/3, 0, 1/3, 1 \right\}, 1\leq j \leq 5 \right\}.
 	\end{eqnarray*}

### 2021-02-18 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
Fix list of references

--- 2021-02-18 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
@@
 
+\bibitem{Strongin2000}
+R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global optimization with non-convex constraints.
+Sequential and parallel algorithms} (Kluwer Academic Publishers, Dordrecht, 2000, 2nd ed.
+2013, 3rd ed.
+2014).
+
+\bibitem{Strongin2013}
+R.~G.~Strongin, V.~P.~Gergel, V.~A.~Grishagin and K.~A.~Barkalov, \textit{Parallel computations for global optimization problems} (Moscow State University, Moscow, 2013) [In Russian].
+
 \bibitem{Evtushenko2009} 
@@
 %\bibitem{Evtushenko2013} Y.~G.~Evtushenko and M.~A.~Posypkin, \textit{A deterministic approach to global box-constrained optimization}, Optim. Lett. \textbf{7}(4), 819--829 (2013).
-
-\bibitem{Strongin2000}
-R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global optimization with non-convex constraints.
-Sequential and parallel algorithms} (Kluwer Academic Publishers, Dordrecht, 2000, 2nd ed.
-2013, 3rd ed.
-2014).
-
-\bibitem{Strongin2013}
-R.~G.~Strongin, V.~P.~Gergel, V.~A.~Grishagin and K.~A.~Barkalov, \textit{Parallel computations for global optimization problems} (Moscow State University, Moscow, 2013) [In Russian].
 
@@
 \bibitem{Gaviano}
-M.~Gaviano, D.~Lera, D.~E.~Kvasov and Ya.~D.~Sergeyev, \textit{Software for generation of classes of test functions with known local and global minima for global optimization}, ACM Trans.
+M.~Gaviano, D.~Lera, D.~E.~Kvasov and Y.~D.~Sergeyev, \textit{Software for generation of classes of test functions with known local and global minima for global optimization}, ACM Trans.
 Math.

### 2021-03-03 2020/JPDC_MCO/2020_JPDC.tex
1. JPDC поправлен список литературы

@@
 \begin{abstract}
-This paper proposes an efficient approach for the parallel solution of computationally time consuming problems of multiple global optimization problems, in which minimized functions can be multiextremal and calculating function values may require huge amounts of computations.
-The proposed approach is based on the information-statistical theory of global optimization, at which a general computational scheme of global optimization methods is proposed.
+This paper proposes an efficient approach for the parallel solution of computationally time consuming problems of multiple global optimization, in which minimized functions can be multiextremal and calculating function values may require huge amounts of computations.
+The proposed approach is based on the information-statistical theory of global optimization, within which a general computational scheme of global optimization methods is proposed.
 In the paper, this general scheme is expanded by the possible reuse of search information obtained in the process of computations when solving multiple global optimization problems.
@@
 \begin{keyword}
-Global optimization\sep dimensionality reduction\sep search information\sep parallel methods of global search\sep computational complexity.
+Global optimization \sep dimensionality reduction \sep search information \sep parallel methods of global search \sep computational complexity.
 \end{keyword}
@@
 
-Global optimization (GO) problems are among the most complex optimization problems.
-In contrast to local optimization problems, when searching for a global minimum, the numerical method must evaluate the behavior of the optimized function in the entire global search area.
+\cite{c1,c2,c3,c4,c5,c6,c7,c8,c9,c10}
+\cite{c11,c12,c13,c14,c15,c16,c17,c18,c19,c20}
+\cite{c21,c22,c23,c24,c25,c26,c27,c28,c29,c30}
+\cite{c31,c32,c33,c34,c35,c36,c37,c38,c39,c40,c41}
+
+Since the early 1960s, the problems of global optimization (GO) have continued to be a grand challenge to the theory and practice of optimization.
+In contrast to local optimization problems, when searching for a global minimum, the numerical method must evaluate the behavior of the function to be optimized in the entire global search domain.
 In general, GO problems are subject to the ``dimensionality curse'', when the volume of computations performed increases exponentially with increasing dimensionality.
 Such computational complexity, as well as high practical demand, have led to a great deal of research activity in the field of global optimization.
-As a result of this research, numerous effective methods for solving global optimization problems and examples of solving practical problems have been proposed -- see, for example, \cite{c1,c2,c3,c4,c5,c6,c7,c8,c9,c10,c11,c12,c13,c14,c18}.
+An overview of approaches developed for solving global optimization problems is presented, for example, in a number of recent monographs and reviews -- see \cite{c1,c2,c3,c4,c5,c6,c7,c8,c9}.
 

### 2021-03-04 2020/JPDC_MCO/2020_JPDC.tex
1.JPDC текст до раздела 4.3

@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+%\usepackage{algorithmic}
+%\usepackage{algorithm}
+\usepackage{algorithm,algpseudocode,float}
+
+\algnewcommand\algorithmicforeach{\textbf{for each}}
+\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
 
@@
 
-\cite{c1,c2,c3,c4,c5,c6,c7,c8,c9,c10}
-\cite{c11,c12,c13,c14,c15,c16,c17,c18,c19,c20}
-\cite{c21,c22,c23,c24,c25,c26,c27,c28,c29,c30}
-\cite{c31,c32,c33,c34,c35,c36,c37,c38,c39,c40,c41}
-
 Since the early 1960s, the problems of global optimization (GO) have continued to be a grand challenge to the theory and practice of optimization.
@@
 
-The paper proposes a further complication in formulating the problem, when not only one, but several global optimization problems, need to be solved simultaneously.
+The first and obvious approach is to use local methods that allow one to efficiently solve optimization problems with unimodal objective functions \cite{c1}.
+In the case of multiextremal problems, local methods should be applied repeatedly to find global minima using different initial points.
+Research in this direction began as early as the 1960s and is still relevant today \cite{c10,c11,c12}.
+Unfortunately, the problem of choosing initial points still remains open -- even using a significant number of initial points does not guarantee a reliable solution of global optimization problems.
+Another obvious approach is to construct in the search domain  a number of grids in whose nodes values of functions to be minimized are computed \cite{c2,c3,c4,c5}.
+To reduce computational complexity of uniform grids, this approach envisages different methods for construction of non-uniform grids, which would be denser only in the vicinity of the global minimum.
+This line of research is developing in the field of so-called nature-inspired metaheuristic methods, including, for example, genetic algorithms, bee algorithms, particle swarm optimization, simulated annealing, ant colony optimization, and firefly algorithms, etc.
+\cite{c4,c13,c14,c15}.
+A large number of applied problems have been solved using these methods, however, as before, it is also difficult to obtain guaranteed estimates of the accuracy when finding globally-optimal solutions for these algorithms.
+
+Yet another line of research in the field of global optimization is based on the assumption that the analytical form of the functions to be optimized is known.
+In this case, we can speak of d.c.(difference of convex) optimization problems in which nonconvex multiextremal functions are represented as a difference of two convex functions \cite{c16,c17}.
+Another approach to the development of global optimization methods in this area is based on the use of interval analysis methods \cite{c18}.
+
+In case analytical expressions for the functions to be optimized are not known (and this is a typical situation for choosing the best solutions in applied optimization problems), it should be understood that numerical estimates of global optimal solutions cannot in principle be obtained if there are no a priori assumptions about the behaviour of the functions to be minimized.
+One if the most widely used ways for specifying such information is to assume the validity of the Lipschitz condition according to which possible changes of the function to be minimized are restricted by the variation values of the parameters being varied.
+The use of this assumption leads to a class of Lipschitzian global optimization problems that has been extensively explored \cite{c3,c6,c7,c19,c19,c20,c21,c22,c23}.
+As a rule, algorithms for solving this class of problems are adaptive, i.e.
+when determining the points for global search iterations, algorithms tend to use all the search information obtained in the process of computation.
+Such adaptivity leads to non-uniform grids and makes it possible to construct estimates of the global minimum based on a finite number of global search iterations.
+A large number of results in this direction have been obtained within the framework of the rapidly developing information-statistical theory of global optimization \cite{c6}.
+It should also be noted that multivariate algorithms of this class quite often use dimensionality reduction using Peano curves or nested optimization scheme \cite{c23}.
+One can also mention the class of DIRECT-type methods using several estimates of the Lipschitz constant chosen from some set of possible values; a detailed review of this line of research is given in \cite{c22}.
+
+The paper proposes a further complication in formulating the decision making problems, when not only one, but several global optimization problems need to be solved simultaneously.
 Such problems are encountered, for example, when one needs to find several efficient solutions of multicriteria optimization problems or when part of the variable parameters in the optimization problem can take only a certain number of discrete values.
-The formulation of such multiple global optimization (MGO) problems is relatively new; for the first time, this type of problem was considered in \cite{c15,c16,c17}.
-The limited attention given to MGO problems is most likely due to the high computational complexity involved.
+For the first time, the formulation of such multiple global optimization (MGO was considered in \cite{c25,c26}, and then was investigated in \cite{c27,c28}.
+The relatively limited attention given to MGO problems is most likely due to the high computational complexity involved for solving even one GO problem.
 
 In general, MGO problems can be solved by a simple sequential (or parallel) solution of individual problems from a set.
-However, this approach means the computational complexity of solving problems increases many times.
-Meanwhile, MGO problems may have a certain informational interrelation -- a similar connectivity is observed, for example, when MGO problems are generated to solve the same multicriteria optimization problem -- see, for example, \cite{c19,c20}.
+However, this approach results in a many-fold increase in the computational complexity of solving such problems.
+Meanwhile, MGO problems may have a certain informational interrelation -- a similar connectivity is observed, for example, when methods based on scalarization of the vector efficiency criterion are applied to solve multicriteria optimization (MCO) problems.
+In this case, it becomes necessary to solve multiple scalar (in the general case, global) optimization problems in order to find several efficient decisions  for MCO problems -- see, for example, \cite{c29,c30,c31,c32,c33}.
 Such interconnectedness of MGO problems can be used to reduce significantly the amount of computation performed when solving the entire set of problems.
 
-The approach proposed in this paper assumes that the computational values of any function of the solvable set of MGO problems can be reduced to the values of any other functions of the set without any time consuming calculations.
+The approach proposed in this paper assumes that the computational values of any function of the set of MGO problems can be recalculated to the values of any other functions of the set without any time consuming calculations.
 In such cases, all the search information obtained in solving a particular problem can be used in solving all other problems in the set.
@@
 In addition, the presence of informational connectivity significantly expands the possibilities of a parallel solution for MGO problems.
-The paper proposes two parallel computational schemes: a cooperative scheme, which means the search information between the problems is exchanged when solving several MGO problems in parallel, and a competitive scheme, which occurs when MGO problems solved in parallel compete with each other for the use of available computational resources.
+The paper proposes two new parallel computational schemes: a cooperative scheme, which means the search information between the problems is exchanged when solving several MGO problems in parallel, and a competitive scheme, which occurs when MGO problems solved in parallel compete with each other for the use of available computational resources.
 
 The structure of the paper is as follows.
-Chapter \ref{sec:2} sets out the problem of multiple global optimization.
-Chapter \ref{sec:3} presents a general computational scheme for solving global search problems.
-Chapter \ref{sec:4} proposes parallel methods for solving MGO problems by using computational systems with shared and distributed memory.
-Chapter \ref{sec:5} contains the results of numerical experiments confirming the effectiveness of the proposed parallel methods.
+Section \ref{sec:2} sets out the problem of multiple global optimization.
+Section \ref{sec:3} presents a general computational scheme for solving global search problems.
+Section \ref{sec:4} proposes parallel methods for solving MGO problems by using computational systems with shared and distributed memory.
+Section \ref{sec:5} contains the results of numerical experiments confirming the effectiveness of the proposed parallel methods.
 In the conclusion, results are discussed and the main directions for further research are presented.
@@
 where $y = (y_1,y_2,\dots,y_N)$ is the vector of varied parameters, $N$ is the dimension of the global optimization problems to be solved, and $D$ is the search domain for the given vectors $a$ and $b$.
+
 In this paper, it is assumed that the functions $f_i(y)$, $1 \leq i \leq s$, can be multiextremal, and the procedures for calculating the values of these functions at points in the search domain $y \in D$ can be computationally expensive.
@@
 
-A key condition of this approach is the assumption that any calculated value of any function $f_i(y)$, $1 \leq i \leq s$, of the set $F$ at an arbitrary point $y \in D$ of the search domain $D$ can be transformed to the value of any other function of the set $F$ at the same point $y \in D$ without any time consuming computations, in a time significantly shorter compared to the computation time values of the functions $f_i(y)$, $1 \leq i \leq s$.
+A key condition of this approach is the assumption that any calculated value of any function $f_i(y)$, $1 \leq i \leq s$, of the set F at an arbitrary point $y \in D$ of the search domain $D$ can be transformed to the value of any other function of the set $F$ at the same point $y \in D$ without any time consuming computations, in a time significantly shorter compared to the computation time values of the functions $f_i(y)$, $1 \leq i \leq s$.
 In other words, there must be a transformation $Z$ such that
-
 \begin{equation}\label{eq:5}
@@
 
-This condition is fulfilled, in part, when solving multicriteria optimization problems (MCO)
-
+This condition is fulfilled, for example, when solving multicriteria optimization (MCO) problems 
 \begin{equation}\label{eq:6}
@@
 
-To find efficient (unimprovable) solutions, a widely used approach is based on the reduction (scalarization) of the vector criterion $g(y)$ to the general scalar criterion, the minimization of which will lead to the Pareto optimal solution of the MCO problem \cite{c19,c20}.
+To find efficient (unimprovable) solutions, a widely used approach is based on the reduction (scalarization) of the vector criterion $g(y)$ to the general scalar criterion, the minimization of which will lead to the Pareto optimal solution of the MCO problem \cite{c29,c30,c31,c32,c33}.
 So, when using mini-max convolution, the vector criterion $g(y)$ from (\ref{eq:6}) reduces to the scalar criterion
@@
 
-If it is necessary to find several efficient decisions to the MCO problem (\ref{eq:6}), the optimization of the scalar criterion $G(\lambda,y)$ from (\ref{eq:7}) is performed for different values of the coefficients $\lambda \in \Lambda$ of the convolution of the criteria $g_i(y)$, $1 \leq i \leq s$.
-Thus, the problem of finding a set of efficient decisions for the MCO problem can be represented as a statement of the MGO problem (\ref{eq:1})-(\ref{eq:3}), where the optimized functions $f_i(y)$, $1 \leq i \leq s$, and the sets $F$ are defined as
+If it is necessary to find several efficient decisions to the MCO problem (\ref{eq:6}), the optimization of the scalar criterion $G(\lambda,y)$ from (\ref{eq:7}) is performed for different values of the coefficients $\lambda \in \Lambda$.
+Thus, the problem of finding a set of efficient decisions for the MCO problem can be represented as a statement of the MGO problem (\ref{eq:1})-(\ref{eq:3}), where the optimized functions $f_i(y)$, $1 \leq i \leq s$, and the set $F$ are defined as
 \begin{equation}\label{eq:8}
@@
 
-Thus, if the values of efficiency criteria $g_i(y)$, $1 \leq i \leq s$, computed at some point $y=\bar{y}$ of the search domain $D$, the value of the scalar criterion $G(\lambda,y)$ (and, accordingly, the values of the functions $f_i(y)$, $1 \leq i \leq s$, from a set $F$) can be computed at the point $\bar{y} \in D$ for any values of the coefficients $\lambda \in \Lambda$ of the convolution $g_i(y)$, $1 \leq i \leq s$ (the transformation of $Z$ from (\ref{eq:5}) is determined by the expression (\ref{eq:7}) at fixed values of the criteria $g_i(y)$, $1 \leq i \leq s$) \cite{c15,c16,c17}.
+As a result, if the values of efficiency criteria $g_i(y)$, $1 \leq i \leq s$, are computed at some point $y=\bar{y}$ of the search domain $D$, then the value of the scalar criterion $G(\lambda,y)$ can be computed at the point $\bar{y} \in D$ for any values of the coefficients $\lambda \in \Lambda$.
+In this case the transformation of $Z$ from (\ref{eq:5}) is determined by the expression (\ref{eq:7}) at fixed values of the criteria $g_l (y)$,$1 \leq l \leq m$) \cite{c27,c28}.
 
-The ability to quickly convert the calculated values of optimized functions determines the need to accumulate and use of search information obtained in the calculation process
-
+The ability to quickly convert the calculated values of optimized functions determines the need to accumulate and use search information obtained in the calculation process
 \begin{equation}\label{eq:9}
@@
 \end{equation}
-where $j$, $1 \leq j \leq s$, is the number of the current minimized function $f_j(y)$ of the set $F$ from (\ref{eq:3}), $k$ is the number of iterations of the global search, $y^i \in D$, $1 \leq i \leq k$,  are points in which the values of the optimized problems of the set $F$ were computed, and $z^i$, $1 \leq i \leq k$, are the values of the function $f_j(y)$ at these points.
+where $j$, $1 \leq j \leq s$, is the number of the current function being minimized $f_j (y)$ of the set $F$ from (\ref{eq:3}), k is the number of iterations of the global search, $y^i \in D$, $1 \leq i \leq k$,  are points in which the values of the optimized problems of the set $F$ were computed, and $f^i$, $1 \leq i \leq k$, are the values of the function $f_j (y)$ at these points (the procedure for calculating the value of the function will be called a trial).
 Global optimization methods can use this search information, thereby ensuring the ability to solve each successive problem of the $F$ set, taking into account the results of all previous calculations.
@@
 
-In conclusion, it should be noted that, in general, the set of minimized functions $F$ from (\ref{eq:3}) can be formed dynamically in the course of calculations by adding new or removing existing global optimization functions
-
+In conclusion, it should be noted that, in general, the set of functions being minimized $F$ from (\ref{eq:3}) can be formed dynamically in the course of calculations by adding new or removing existing global optimization functions
 \begin{equation}\label{eq:10}
@@
 
-The approach proposed in this paper is based on two key ideas used in the framework of the information-statistical theory of global optimization \cite{c5}.
+The approach proposed in this paper is based on two key ideas used in the framework of the information-statistical theory of global optimization \cite{c6}.
 
-In accordance with this theory, first of all, dimension reduction is achieved using the Peano space-filling curves (evolvents) $y(x)$, uniquely and continuously mapping the interval $[0,1]$ to the $N$-dimensional domain $D$ -- see, for example, \cite{c5,c21}.
-As a result of this reduction, multidimensional global optimization problems (\ref{eq:7}) can be reduced to one-dimensional ones:
+In accordance with this theory, first of all, dimensionality reduction is achieved using the Peano space-filling curves (evolvents) $y(x)$, uniquely and continuously mapping the interval [0,1] to the $N$-dimensional domain $D$ -- see, for example, \cite{c6,c23}.
+As a result of this reduction, multidimensional global optimization problems (\ref{eq:1})-(\ref{eq:2}) can be reduced to one-dimensional ones:
 
@@
 After dimensionality reduction, the search information $\omega_k$ from (\ref{eq:9}) takes the form
-
 \begin{equation}\label{eq:13}
@@
 \end{equation}
-for a more efficient execution of global search algorithms.
+or a more efficient execution of global search algorithms.
 It should be noted that in optimizing the reduced one-dimensional functions $f_i (y(x))$, $1 \leq i \leq s$, many one-dimensional global search algorithms can be used (possibly after some generalization).
 
-Numerical methods for constructing approximations of Peano space-filling curves (evolvents) with a given accuracy for a given dimensionality are considered in \cite{c5}; examples of such approximations are shown in Fig.
-\ref{fig:1}.
+Numerical methods for constructing approximations of Peano space-filling curves (evolvents) with a given accuracy for a given dimensionality are considered in \cite{c6}; examples of such approximations are shown in Fig.\ref{fig:1}.
+
+Further, in the framework of information-statistical theory, a general computational scheme of global optimization algorithms is proposed \cite{c6}. This scheme can be summarized as follows.
 
@@
 
-Further, in the framework of information-statistical theory, a general computational scheme of global optimization algorithms is proposed. This scheme can be summarized as follows.
-
-Let $k$, $k \geq 2$, optimization iterations be performed, and the search information $A_k$ from (\ref{eq:13}) is obtained.
-To adaptively select the points of the next iterations, the optimization algorithm has to evaluate the possibility of placing the global minimum in the subintervals, into which the interval $[0,1]$ is divided by the points $x_i$, $1 \leq i \leq k$, previously performed iterations of the global search.
-This assessment can be performed by introducing characteristics of the subintervals, the value of which would be proportional to the degree to which the global minimum can be located in these subintervals
-
+Let $f(y)$ be the function to be minimized for which $k$, $k \geq 2$, optimization iterations are performed, and the search information $A_k$ from (\ref{eq:13}) is obtained.
+To adaptively select the points of the next iterations, the optimization algorithm has to evaluate the possibility of placing the global minimum in the subintervals, into which the interval [0,1] is divided by the points $x_i$, $1 \leq i \leq k$, of previously performed iterations of the global search.
+This assessment can be performed by introducing \textit{characteristics} of the subintervals, the value of which would be proportional to the degree to which the global minimum can be located in these subintervals
 \begin{equation}\label{eq:15}
@@
 For example, for an algorithm that builds a uniform dense grid in the search domain, the characteristics may simply be the length of the interval
+\begin{equation}\label{eq:16}
+R(i)=(x_i-x_{i-1}), 1 < i \leq k,
+\end{equation}
+where the points $x_i$, $1 < i \leq k$, are from $A_k$.
 
-\begin{equation}\label{eq:16}
-R(i)=(x_i-x_{i-1}), 1 < i \leq k, \; x_i, x_{i-1} \; \mathrm{from} \; A_k.
-\end{equation}
-
-For the multidimensional generalizations of the algorithms proposed in \cite{c22,c23}, the characteristic is an estimate of the minimum possible value of the minimized function on the interval, i.e.
-
+For the multidimensional generalizations of the algorithms proposed in \cite{c34,c35}, the characteristic is an estimate of the minimum possible value of the function being minimized on the interval, i.e.
 \begin{equation}\label{eq:17}
@@
 \end{equation}
-where $H$ is the H\"older constant from (\ref{eq:12}) for the reduced global optimization problem to be solved; $N$ is the dimensionality of the problem from (\ref{eq:7}).
-For the multidimensional generalized global search algorithm (GSA) \cite{c5,c24}, developed in the framework of the information-statistical approach, the characteristic has the form
-
+where $H$ the values $z_i$, $1<i \leq k$, are from $A_k$, H is the H\"older constant from (\ref{eq:12}) for the reduced global optimization problem to be solved and $N$ is the dimensionality of the problem from (\ref{eq:1}).
+For the multidimensional generalized global search algorithm (GSA) \cite{c6,c23}, developed in the framework of the information-statistical approach, the characteristic has the form
 \begin{equation}\label{eq:18}
@@
 \end{equation}
-($r>1$ is the parameter of the GSA algorithm).
+($r>1$ is the reliability parameter of the GSA algorithm).
 
-The presence of interval characteristics makes it possible to represent the procedure for iterating a global search as the following sequence of steps \cite{c5}.
+A detailed description of the GSA algorithm is given in Algorithm \ref{alg:1}(see also \cite{c6}).
+The inputs of the algorithm are the function to be minimized $f(y)$\footnote{ }, the search domain $D$ from (\ref{eq:2}), the reliability parameter $r > 1$, the required accuracy $\varepsilon > 0$ and the maximum permissible number of global search iterations $kmax$.
+The output of the algorithm is an estimate of the global minimum point $y_k^* =y(x_k^*)$, the value of the function to be minimized at this point $z_k^* =f(y_k^*)$  and the number of performed global search iterations $k$.
+In the course of calculations, the search information $A_k$\footnote{ } from (\ref{eq:13}) is used, which contains the points of previously performed global search iterations $x_i$, $1 \leq i \leq k$, ordered in ascending order, and the values of the function to be minimized $z_i=f(y(x_i))$, $1\leq i \leq k$.
 
-\textit{Step 1.} Calculate the characteristics of the intervals $R(i)$, $1 < i \leq k$, and determine the interval with the maximum characteristic
+\begin{algorithm}[]
+\caption{PSEUDO CODE OF THE GSA ALGORITHM} \label{alg:1}
+\scriptsize
+\hspace*{\algorithmicindent} \textbf{Input}: $f(y)$, $D$, $r$, $\varepsilon$, $K$;\\
+\hspace*{\algorithmicindent} \textbf{Output}: $y_k^*$, $z_k^*$, $k$;
+\begin{algorithmic}[1]
+\State Initialize $A_2=\{(x_1=0,y_1=y(x_1 ),z_1=f(y(x_1 )),(x_2=1,y_2=y(x_2 ),z_2=f(y(x_2 ))\}$, $k = 2$, $stop = false$;
+\While{$k < kmax$ \textbf{and not} $stop$} 
+  \State Calculate the estimate of the H\"older constant $m$ from (\ref{eq:19}) using $A_k$;
+  \State Calculate the interval characteristics $R(i)$ from (\ref{eq:18}) \textbf{for all} $i = 2, \dots, k$;
+	\State Find the maximum characteristic 
 	\begin{equation}\label{eq:20}
@@
   \end{equation}
-
-\textit{Step 2.} Select the next iteration point in the interval with the maximum characteristic \footnote{The rule $X$ for choosing the point of the next iteration in the interval $(x_{t-1},x_t)$ is another design element of characteristically represented algorithms.}
+	\State Calculate the next iteration point
 	\begin{equation}\label{eq:21}
-x^{k+1}=X(x_{t-1},x_t ).
+	\begin{matrix}
+	   x^{k+1}=\frac{x_t+x_{t-1}}{2}-sign(z_t-z_{t-1})(\frac{|z_t-z_{t-1}|}{m})^N/2r \\
+		 sign(v)=\begin{cases}
+		            1, &v>0,\\
+		            0, &v=0,\\
+								-1,&v<0
+							\end{cases}
+	\end{matrix}
 	\end{equation}
 
-Calculate the value $z^{k+1}$ of the minimized function at this point (the procedure for calculating the value of the function will be called a trial) and add the obtained data $(x^{k+1},z^{k+1})$ to the search information $A_k$ from (\ref{eq:13}).
+	\State
+	\begin{equation}\label{eq:22}
+     stop = (x_t - x_{t-1})^{1/N} \leq \varepsilon.
+  \end{equation}
+  \If{ \textbf{not} $stop$}
+	  \State Calculate $z^{k+1} = f( y(x^{k+1}))$;
+    \State Insert $(x^{k+1}, y(x^{k+1}), z^{k+1})$ into the search information $A_k$;
+    \State Increase $k = k + 1$;
+  \EndIf
+\EndWhile
+\State Find the estimate of the global minimum point
+\begin{equation}\label{eq:23}
+y_k^*= y(x_l),  z_k^*=z_l,
+\end{equation}
+where $l= \argmin_{1 \leq i \leq k}z_i$.
+\State \textbf{return} $y_k^*$, $z_k^*$, $k$;
+\end{algorithmic}
+\end{algorithm}
 
-\textit{Step 3.} Check the stopping condition 
-\begin{equation}\label{eq:22}
-\rho_t \leq \varepsilon,	
-\end{equation}
-where $\rho_t=(x_t-x_{t-1})^{1/N}$, $t$ from (\ref{eq:20}) and $\varepsilon > 0$ is the specified accuracy of the solution to the optimization problem.
-If the stopping condition (\ref{eq:22}) is reached, then the solution to the optimization problem is stopped, otherwise $k=k+1$ is assumed and the next iteration of the global search is performed.
-
-After the calculations are completed, the lowest calculated value of the minimized function can be taken as the global minimum estimate
-\begin{equation}\label{eq:23}
-z_k^*=\min{z_i}, 1 \leq i \leq k.
-\end{equation}
-
-Again, we note that the above computational scheme is quite general.
-Within the framework of this scheme, many global search algorithms can be presented, as evidenced, in particular, by the examples given in (\ref{eq:16})--(\ref{eq:18}), as well as other characteristically represented algorithms (see, for example, \cite{c22,c23,c24,c25,c26,c27,c28,c29,c30,c31}).
-Thus, the parallelization methods discussed in Section \ref{sec:4} can be used for a sufficiently wide variety of global search algorithms.
-
-The convergence conditions for the characteristically represented algorithms depend on the properties of the interval characteristics used.
-For example, one of the sufficient conditions for convergence of the algorithms is the requirement that, when iterating the global search, the characteristic of the interval containing the global minimum point repeatedly takes the maximum value in Step 1 of the general characteristic scheme.
-This condition is fulfilled, for example, for the multidimensional generalizations of the algorithms proposed in \cite{c22,c23} for the exact setting of the H\"older constant in (\ref{eq:17}).
-For GSA, the sufficient condition for convergence is the relation \cite{c5}
+The sufficient condition for convergence of the GSA algorithm is the relation
 \begin{equation}\label{eq:24}
@@
 \end{equation}
-starting with some iteration of the global search. Additional information on the convergence of characteristically represented global search algorithms can be obtained, for example, in \cite{c5}.
+that should be satisfied starting with some iteration of the global search.
 
-In the framework of the proposed approach, the GSA algorithm described by rules (\ref{eq:18})-(\ref{eq:23}) will be used to solve global optimization problems of the set $F$.
-Moreover, due to the informational connectivity (\ref{eq:5}) of the problems of the set $F$, the GSA algorithm can use all the search information $A_k$ from (\ref{eq:13}) when solving the next global optimization problem.
+In the framework of the proposed approach, the GSA algorithm described above will be used to solve global optimization problems of the set $F$.
+Moreover, due to the informational connectivity (\ref{eq:5}) of the problems of the set $F$, the search information $A_k$ can be quickly transformed to the values of the next function to be minimized $f(y)$ of the set $F$ and GSA can minimize each successive function $f(y)$, taking into account the results of all previous calculations.
 The GSA algorithm for solving problems of the set $F$ and using search information $A_k$, will be referred to as the multiple global search algorithm (MGSA).
@@
 
-In keeping with the initial assumption that it is computationally expensive to calculate the values of minimized functions of the set $F$ from (\ref{eq:3}), solving all global optimization problems (\ref{eq:1}) may require a huge amount of calculations, which increases with the number of problems to be solved.
+In keeping with the initial assumption that it is computationally expensive to calculate the values of the functions to be minimized of the set $F$ from (\ref{eq:3}), solving all global optimization problems (\ref{eq:1}) may require a huge amount of calculations, which increases with the number of problems to be solved.
 Thus, obtaining estimates of globally optimal decisions with reasonable time delays can only be achieved using parallel high-performance computational systems.
 
-An obvious way to apply parallel computing is to solve the problems of the set $F$ in parallel.
-This method is simple and does not require any additional efforts for implementation.
-With this approach and with a sufficient number of computational devices (cores / processors), the time needed to solve all the $F$ set will be determined by the time used in solving the most time-consuming problem.
+An obvious way to apply parallel computing is to solve the problems of the set F in parallel.
+Undoubtedly this approach is straightforward and does not require any additional efforts to be  implemented.
+In this case, if a sufficient number of computational devices (cores / processors) is available, the time needed to solve all problems from the F set will be determined by the time used in solving the most time-consuming problem.
 However, this time may turn out to be quite long and necessitate the use of parallel computing to solve even one global optimization problem.
+However, well-known and widely used parallelization methods are poorly applicable in the case of global optimization problems.
+For example, the search domain distribution between processors for global optimization problems leads to the global minimum of the problems to be solved being located on only one of the processors, and all other processors will perform redundant calculations.
 
-Unfortunately, well-known and widely used parallelization methods are poorly applicable in the case of global optimization problems.
-For example, search domain distribution between processors for global optimization problems leads to the global minimum of the problems to be solved being located on only one of the processors; in this case, all other processors will perform redundant calculations.
+In the framework of the information-statistical theory, a general approach to parallelizing global optimization problems has been proposed -- parallel computing is provided by parallel (simultaneous) computing of the values of the minimized function at several different points in the search domain.
+This approach is general -- it can be applied to solving any global optimization problems in a parallel mode.
+And such an approach is efficient, since the most time consuming part of the problem is parallelized due to the initial assumption that calculating the values of functions to be minimized is computationally expensive.
 
-In the framework of the information-statistical theory, a general approach to parallelizing global optimization problems has been proposed -- parallel computing is provided by parallel (simultaneous) computing of the values of the minimized function at several different points in the search domain \cite{c5}.
-This approach is general -- it can be applied to solving any global optimization problems in a parallel mode.
-And such an approach is efficient, since the most time consuming part of the problem is parallelized due to the initial assumption that calculating the values of minimized functions is computationally expensive.
-
-Next, we propose three different ways of applying this approach for parallel solutions to multiply global optimization problems (\ref{eq:1})-(\ref{eq:3}).
+Next, we propose three different methods of applying this approach for parallel solutions to multiply global optimization problems (\ref{eq:1})-(\ref{eq:3}).
 
@@
 
-In this method of parallel computing, it is proposed to solve the set $F$ problems strictly sequentially, and to use all available computational resources only to solve the one optimization problem.
+In this method of parallel computing, it is proposed to solve the set $F$ of problems strictly sequentially, and to use all available computational resources only to solve just the optimization problem.
 When solving problems of the set $F$ sequentially, all available search information $A_k$, obtained in the calculation process is reused.
 
-To implement this approach, it is necessary to ensure that at each iteration of the global optimization problem, not one, but several points in the search domain $D$ are selected for parallel calculation of the values of the minimized function at these points.
-In keeping with the fact that characteristically represented global search algorithms use the characteristics of the intervals $R(i)$, $1 < i \leq k$, from (\ref{eq:15}), to assess the possibility of locating the global minimum in the subinterval of the interval $[0,1]$, it is most reasonable to select the points of the next iteration of the global search in the intervals with the maximum values of the characteristics \cite{c33,c34,c35}.
+To implement this approach, it is necessary to ensure that at each iteration of the global optimization problem, not one, but several points in the search domain $D$ are selected for parallel calculation of the values of the functions to be minimized at these points.
+In keeping with the fact that characteristically represented global search algorithms use the characteristics of the intervals $R(i)$, $1 < i \leq k$, from (\ref{eq:15}), to assess the possibility of locating the global minimum in the subinterval of the interval [0,1], it is most reasonable to select the points of the next iteration of the global search in the intervals with the maximum values of the characteristics \cite{c6}.
 
-Performing iterations of a global search for a parallel generalization of the MGSA algorithm (PMGSA-1) can be represented as follows ($q>1$ is the number of available computational devices (cores/processors) with shared memory) \cite{c33,c34,c32}.
+A parallel generalization of the MGSA algorithm (PMGSA-1) is given in Algorithm \ref{alg:2}.
 
-\textit{Step 1}. Calculate the characteristics of the intervals $R(i)$, $1 < i \leq k$, and determine $q$ intervals with the maximum values of characteristics
+\begin{algorithm}[]
+\caption{PSEUDO CODE OF THE PMGSA-1 ALGORITHM} \label{alg:2}
+\scriptsize
+\hspace*{\algorithmicindent} \textbf{Input}: $F(y)$, $D$, $r$, $\varepsilon$, $kmax$, $q$;\\
+\hspace*{\algorithmicindent} \textbf{Output}: $Y$, $Z$, $k$;
+\begin{algorithmic}[1]
+\State Initialize $ k=2, A_2=\{(x_1=0,y_1=y(x_1),z_1=f(y(x_1)),(x_2=1,y_2=y(x_2),z_2=f(y(x_2))\} $;
+\State $Y = \{\emptyset \}, Z = \{\emptyset\}$
+\State Set $stop = false$;
+\ForEach {$f(y)\subset \mathcal{F}(y)$ \textbf{and not} $stop$ }
+\While{$k < kmax$ \textbf{and not} $stop$} 
+  \State Calculate the estimate of the H\"older constant $m$ from (\ref{eq:19}) using $A_k$;
+  \State Calculate the interval characteristics $R(i)$ from (\ref{eq:18}) \textbf{for all} $i = 2, \dots, k$;
+	\State Set $T = \{ \emptyset \}, XK = \{ \emptyset \}$;
+	\ForEach { $i = 1..q$ } 
+	// q is the number of available cores/processors
+	  \State Find the interval with the maximum characteristic 
+		\begin{equation*}
+			t= \argmax_{1<i \leq k} R(i), i \notin T;
+	  \end{equation*}
+		\State Calculate the iteration point $x^{k+i}$ in accordance with (\ref{eq:21});
+    \State Insert $t$ in $T$ and $x^{k+i}$ in $XK$;
 
-\begin{equation}\label{eq:25}
-R(t_j)=\max{R(i)},1 < i \leq k, 1 \leq j\leq q,
-\end{equation}
-(for $j>1$,$t_j \neq t_l$, $1 \leq l < j$ should be satisfied).
+	\EndFor
+  \State \textbf{for each} $t \in T$ \textbf{do if} $(x_t-x_{t-1} )^{1/N} \leq \varepsilon$ \textbf{then} $stop = true$;
 
-\textit{Step 2}. Select $q$ points of the next parallel iteration in the intervals with maximum characteristics
-
-\begin{equation}\label{eq:26}
-x^{k+j}=X(x_{t_j-1},x_{t_j} ), 1 \leq j \leq q,
-\end{equation}
-then calculate the values of $z^{k+j}$, $1 \leq j \leq q$, of the minimized function at these points and add the resulting data $(x^{k+j},z^{k+j})$, $1 \leq j \leq q$, to the search information $A_k$ from (\ref{eq:13}).
-
-\textit{Step 3}. Check the stopping condition
-
-\begin{equation}\label{eq:27}
-\rho_{t_j} \leq \varepsilon, 1 \leq j \leq q,
-\end{equation}
-where $\rho_{t_j}=(x_{t_j}-x_{t_j-1} )^{1/N}$, $t_j$, $1 \leq j \leq q$, from (\ref{eq:25}) and $\varepsilon > 0$ is the specified accuracy for solving the optimization problem.
-If the stopping condition (\ref{eq:27}) is reached for at least one interval $t_j$, $1 \leq j \leq q$, then the solution to the optimization problem stops; otherwise $k=k+q$ is assumed and the next parallel iteration of the global search is performed.
+	\If{ \textbf{not} $stop$}
+	  \ForEach { $x \in XK $ } \textbf{in parallel}
+	    \State Calculate $z = f( y(x))$;
+      \State Insert $(x, y(x), z)$ into the search information $A_k$;
+      \State Increase $k = k + 1$;
+	  \EndFor
+  \EndIf
+	\State Find the estimate of the global minimum point of the current function $f(y)$ in accordance with (\ref{eq:23});
+	\State Insert $y_k^*$ in $Y$ and $z_k^*$ in $Z$;
+  \EndWhile
+\EndFor
+\State \textbf{return} $Y$, $Z$, $k$;
+\end{algorithmic}
+\end{algorithm}
 
@@
 On the left in Fig.
-\ref{fig:2} the MGSA result is given (55 iterations), on the right the PMGSA-1 result is shown (30 iterations, i.e., speedup is 1.8).
+\ref{fig:2} the MGSA result is given (55 iterations), on the right the PMGSA-1 result is shown (30 iterations, i.e.
+the speedup is 1.8).
 PMGSA-1 uses 2 cores, iteration points selected by the cores are marked by the different lines.
@@
   \includegraphics[width=\linewidth]{fig2}
-  \caption{An example of using PMGSA-1 for solving the univariate optimization problem: a) the graph of the minimized function, b) step-by-step selection of iteration points computed by the method (the ordinate axis is the iteration number, the abscissa axis is the search domain)}
+  \caption{An example of using PMGSA-1 for solving the univariate optimization problem: a) the graph of the function being minimized, b) step-by-step selection of iteration points computed by the methods (the ordinate axis is the iteration number, the abscissa axis is the search domain)}
   \label{fig:2}
@@
 
-Selecting several points at once for the same search iteration in the PMGSA-1 algorithm can result in redundant points compared to the sequential GSA version, since these points are selected at the same state of search information $A_k$ from (\ref{eq:13}).
+Selecting several points at once for the same search iteration in the PMGSA-1 algorithm can result in redundant points compared to the sequential MGSA version, since these points are selected at the same state of search information $A_k$ from (\ref{eq:13}).
 Such an undesirable effect can be reduced by reducing the number of computational devices allocated to solve an optimization problem, while simultaneously increasing the number of parallel global optimization problems that are solved due to freed computational resources.
@@
 
-Assume, as before, that $q>1$ is the number of available computational devices (cores/processors) with shared memory, and $ns>1$ is the number of parallel problems solved from the set $F$ from (\ref{eq:3})\footnote{Without lost of generality, we will use sequential numbering of problems solved in parallel from $1$ to $ns$.}.
-Each optimization problem solved in parallel will compete for available computational resources -- the distribution of resources between problems can be carried out, as before, based on the values of interval characteristics.
-To calculate the characteristics, the problems solved in parallel have sets of search information $A_k(j)$, $1\leq j \leq ns$, from (\ref{eq:13}), which match the set of iteration points $x_i$, $1 \leq i \leq k$, of the global search, but differ, accordingly, in the values of the minimized functions.
-The notation $\tau_{ij}$ will then designate the interval $t_j$ in the search information $A_k(i)$ of problem $i$.
+Assume, as before, that $q>1$ is the number of available computational devices (cores/processors) with shared memory, and $ns>1$ is the number of optimization problems\footnote{Without lost of generality, we will use sequential numbering of problems solved in parallel from 1 to $ns$.}  solved in parallel.
+Each optimization problem will compete for available computational resources -- the distribution of resources between problems can be carried out, as before, based on the values of interval characteristics.
+To calculate the characteristics, the problems solved in parallel have their own separate sets of search information $A_k (is)$, $1 \leq is \leq ns$, from (\ref{eq:13}), which match the set of iteration points $x_i$, $1 \leq i \leq k$, of the global search, but differ, accordingly, in the values $z_i (is)$, $1 \leq is \leq k$, of the minimized functions $f_is (y)$, $1 \leq is \leq ns$, from the set $F(y)$.
+As a result, in this case, to specify the intervals where the characteristics are computed, two indices $( is, t )$ are required, where $is$, $1 \leq is \leq ns$, is the number of the function to be minimized  $f_is (y)$ and  $t$, $1 \leq t \leq k$, is the number of the interval in the search information $A_k (is)$.
 
-The new parallel version of the MGSA algorithm (the PMGSA-2 algorithm) obtained as a result of this approach is determined by the following sequence of steps for performing parallel iterations of the global search.
+It should also be noted that when implementing this approach to compare the interval characteristics in different global optimization problems, the values of the function being minimized in the available search information should be normalized\footnote{Without loss of generality, we will use sequential numbering of problems solved in parallel from 1 to $ns$.}
+\begin{equation}\label{eq:25}
+z(is)_i'=\frac{z(is)_i - z(is)_k^*}{m(is)_k}, 1 \leq i \leq k,1 \leq is \leq ns,
+\end{equation}
+where $z(is)_k$, $1 \leq is \leq ns$, are the estimates (\ref{eq:23}) of the minimum values of the functions being minimized, and $m(is)_k$, $1 \leq is \leq ns$, are the estimates (\ref{eq:19}) of the H\"older constants $H_{is}$, $1 \leq is \leq ns$, from (\ref{eq:12}).
+Further, when performing the steps of the algorithm, the converted values $z(is)_i'$, $1 \leq i \leq k$, $1 \leq is \leq ns$, should be used when calculating the characteristics of intervals and points of new iterations of the global search.
 
-\textit{Step 0}.
-For each problem to be solved in parallel, normalize the values of the minimized function in the available search information in order to compare the interval characteristics in different global optimization problems\footnote{After transformation (\ref{eq:28}), $z_k^*=0$ and $m_k=1$ are carried out for all $ns$ parallel problems.}
+The new parallel version of the MGSA algorithm (the PMGSA-2 algorithm) obtained as a result of this approach is given in Algorithm \ref{alg:3}.
 
-\begin{equation}\label{eq:28}
-z(j)_i'=\frac{z(j)_i-z(j)_k^*}{m(j)_k},1 \leq i \leq k,1 \leq j \leq ns,
-\end{equation}
-where $z(j)_k^*$, $1 \leq j \leq ns$, are estimates (\ref{eq:23}) of the minimum values of the minimized functions, and $m(j)_k$, $1 \leq j \leq ns$, are the estimates (\ref{eq:19}) of the H\"older constant from (\ref{eq:12}).
-Further, when performing the steps of the algorithm, the converted values $z(j)_i'$, $1 \leq i \leq k$, $1 \leq j \leq ns$ should be used when calculating the characteristics of intervals and points of new iterations of the global search.
+\begin{algorithm}[]
+\caption{PSEUDO CODE OF THE PMGSA-2 ALGORITHM} \label{alg:3}
+\scriptsize
+\hspace*{\algorithmicindent} \textbf{Input}: $F(y)$, $D$, $r$, $\varepsilon$, $kmax$, $ns$;\\
+\hspace*{\algorithmicindent} \textbf{Output}: $Y$, $Z$, $k$;
+\begin{algorithmic}[1]
+\State Initialize $ k=2, A_2=\{(x_1=0,y_1=y(x_1),z_1=f(y(x_1)),(x_2=1,y_2=y(x_2),z_2=f(y(x_2))\} $;
+\State Set $stop = false$;
+\While{$k < kmax$ \textbf{and not} $stop$}
+  \ForEach { $is = 1..ns$ }
+	  \State Calculate the estimate of the H\"older constant $m$ from (\ref{eq:19}) using $A_k$;
+		\State Find the estimate of the global minimum value $z(is)_k^*$ of the function $f_{is}(y)$ in accordance with (\ref{eq:23});
+		\State Set $A_k'(is)=A_k (is)$;
+		\State Normalize the values $z(is)_i'$, $1 \leq i \leq k$, in $A_k'(is)$ in accordance with (\ref{eq:25});
+		\State Set $m_k (is)=1$;
+	\EndFor
 
-\textit{Step 1}.
-Calculate the characteristics of intervals $R(i)$, $1 < i \leq k$, for all $ns$ problems to be solved in parallel and determine $q$ intervals with the maximum values of the characteristics
-\begin{equation}\label{eq:29}
-R(\tau_{l(j)j})=\max{R_\nu(i)}, 1 < i \leq k, 1 \leq \nu \leq ns, 1 \leq j \leq q,
-\end{equation}
-where $R_\nu(i)$ is the characteristic of the interval $i$ in the search information $A_k(\nu)$ of the problem with the number $\nu$, $l(j)$ is the number of the problem to be solved, in which the maximum characteristic $R(\tau_{l(j)j})$ was chosen; (in (\ref{eq:29}) for $j>1$, $\tau_{l(j)j}\neq \tau_{l(l)l}$, $1 \leq l < j$ should be satisfied).
+	\State Set $T = \{ \emptyset \}, XK = \{ \emptyset \}$;
+	\ForEach { $i = 1..q$ } 
+	// q is the number of available cores/processors
+	  \State Find the interval with the maximum characteristic 
+		\begin{equation*}
+			R_{is} (t)= \argmax_{1 \leq is \leq ns, 1 < i \leq k} R(i),i \notin T 
+	  \end{equation*}
+		using $A_k' (is)$;
+		\State Calculate the iteration point $x^{k+i}$ using $A_k' (is)$ in accordance with (\ref{eq:21});
+    \State Insert $t$ in $T$, is in $IS$ and $x^{k+i}$ in $XK$;
 
-\textit{Step 2}. The execution of Step 2 is divided into three sequential stages: 
-\begin{enumerate}
+	\EndFor
+  \State \textbf{for each} $t \in T$ \textbf{do if} $(x_t-x_{t-1} )^{1/N} \leq \varepsilon$ \textbf{then} $stop = true$;
 
-	\item Select $q$ points of the next parallel iteration in intervals with maximum characteristics
-\begin{equation}\label{eq:30}
- x(l(j))^{k+j}=X(x_{\tau_{l(j)j}-1},x_{\tau_{l(j)j}}),1 \leq j \leq q	
-\end{equation}
-(the point $x(l(j))^{k+j}$ is determined in the interval $(x_{\tau_{l(j)j}-1},x_{\tau_{l(j)j}})$, belonging to the search information $A_k (l(j))$ of problem $l(j)$.
-
-	\item Calculate the values $z(l(j))^{k+j}$, $1 \leq j \leq q$, of minimized functions at the points $x(l(j))^{k+j}$, $1 \leq j \leq q$.
-
-  \item Add the obtained data $(x(l(j))^{k+j}, z(l(j))^{k+j})$, $1 \leq j \leq q$, in the search information $A_k(i)$, $1 \leq i \leq ns$, of each problem being solved in parallel with the necessary values transformation $z(l(j))^{k+j}$, $1 \leq j \leq q$, using the $Z$ transformation from (\ref{eq:5}).
-
-\end{enumerate}
-
-\textit{Step 3}. Verify the stopping condition
-\begin{equation}\label{eq:31}
-\rho_{\tau_{l(j)j}} \leq \varepsilon, 1 \leq j \leq q,
-\end{equation}
-where $\rho_{\tau_{l(j)j}} = (x_{\tau_{l(j)j}} - x_{\tau_{l(j)j)-1}} )^{1/N}$, $\tau_{l(j)j}$, $1 \leq j \leq q$, from (\ref{eq:29}) and $\varepsilon > 0$ is the given accuracy for solving the optimization problem.
-If the stopping condition (\ref{eq:31}) is reached for at least one interval $\tau_{l(j)j}$, $1 \leq j \leq q$, then the solution to the optimization problem stops; otherwise, $k=k+q$ is assumed and the next parallel iteration of the global search is performed.
+	\If{ \textbf{not} $stop$}
+	  \ForEach { $x \in XK $ } \textbf{in parallel}
+		  \State Take $is$ from $IS$;
+	    \State Calculate $z = f_{is}(y(x))$;
+      \State Insert $(x, y(x), z)$ into the search information $A_k$;
+      \State Increase $k = k + 1$;
+	  \EndFor
+  \EndIf
+	\State Set $Y = \{ \emptyset \}, Z = \{ \emptyset \}$;
+	\ForEach { $iz = 1..ns$ }
+	\State Find the estimate of the global minimum point $y_k^*$ using $A_k (is)$ in accordance with (\ref{eq:23});
+	\State Insert $y_k^*$ in $Y$ and $z_k^*$ in $Z$;
+  \EndFor
+	\EndWhile
+\State \textbf{return} $Y$, $Z$, $k$;
+\end{algorithmic}
+\end{algorithm}
 
 Let us briefly explain the rules for performing a parallel iteration of the global search.
-The characteristics of the intervals are calculated for all $ns$ problems to be solved in parallel (Step 1), and $q$ intervals with the maximum values of the characteristics are selected among them.
+The characteristics of the intervals are calculated for all ns problems to be solved in parallel, and q intervals with the maximum values of the characteristics are selected among them (lines 12-16).
 As a result, the available computational resources are distributed dynamically among problems -- computational devices can be distributed evenly between problems, or they can be allocated to only one problem with the best interval characteristics.
-After calculating the values of the minimized functions (Step 2), the new search data obtained is added to the search information of all solved problems; regardless of where the intervals for iterating the global search were selected, the search information of all solved problems is updated with a full set of new search data.
+After calculating the values of the functions being minimized, the new search data obtained is added to the search information of all solved problems; regardless of where the intervals for iterating the global search were selected, the search information of all solved problems is updated with a full set of new search data (lines 19-24).
 
 The maximum efficiency of the proposed PMGSA-2 algorithm is achieved if the global minimum in parallel problems is reached at the same point in the search domain.
-In any case, getting additional search information from simultaneously solved problems can significantly reduce the time needed to solve each global search problem -- see the results of numerical experiments in Section \ref{sec:5}.
-An example how PMGSA-2 selects the intervals with the maximum characteristics (Step 1) is given in Fig.\ref{fig:3}.
+In any case, getting additional search information from simultaneously solved problems can significantly reduce the time needed to solve each global search problem.
+This conclusion is rather obvious if the global minima points of the functions $f_i (y)$, $1 \leq i \leq s$, of the set $F$ from (\ref{eq:3}) are situated in the same small neighborhood of the search domain $D$ from (\ref{eq:2}).
+But even in the case when the global minima points of different functions differ much from each other, the values of different functions $f_i (y)$, $1 \leq i \leq s$, of the set $F$ computed in the process of global search can be used for each function separately.
+Thus, in PMGSA-2 the execution of one parallel iteration using $q$ computational units adds $q$ computation results.
+Increasing the amount of available search information allows adaptive global search algorithms to select points for the next iterations of the global search in a more reasonable manner.
+In the limiting case, when performing sufficiently many $iterations \gg 1$, the amount of search information can be sufficient to find the global minimum point of the next problem in the set $F$ without any additional iterations.
+This effect is proved also by the results of numerical experiments in Section \ref{sec:5}.
+An example of how PMGSA-2 selects the intervals with the maximum characteristics is given in Fig.\ref{fig:3}.
 

### 2021-03-05 2020/JPDC_MCO/2020_JPDC.tex
1. Новый вариант статьи переведен в tex полностью

@@
 \begin{equation}\label{eq:6}
-g(y) = (g_1(y), g_2(y), \dots , g_s(y)) \to min,  y\in D.
+g(y) = (g_1(y), g_2(y), \dots , g_m(y)) \to min,  y\in D.
 \end{equation}
@@
 \begin{matrix}
-G(\lambda,y)=\max{(\lambda_i g_i(y),1 \leq i \leq s)},	\\
-\lambda=(\lambda_1,\lambda_2,\dots,\lambda_s)\in \Lambda \subset R^s:\sum_{i=1}^s{\lambda_i =1},\lambda_i\geq 0,1 \leq i \leq s.
+G(\lambda,y)=\max{(\lambda_l g_l(y),1 \leq l \leq m)},	\\
+\lambda=(\lambda_1,\lambda_2,\dots,\lambda_m)\in \Lambda \subset R^m:\sum_{l=1}^m{\lambda_l =1},\lambda_l\geq 0,1 \leq l \leq m.
 \end{matrix}
@@
 \begin{equation}\label{eq:8}
-f_i (y)= G(\lambda_i,y),1 \leq i \leq s.
+F(y)= \{ G(\lambda_i,y),1 \leq l \leq s\}.
 \end{equation}
@@
 \begin{equation}\label{eq:12}
-|f_i (y(x'))-f_i (y(x''))| \leq H_i \|x'-x''\|, x',x''\in [0,1], 1 \leq i \leq s,
+|f_i (y(x'))-f_i (y(x''))| \leq H_i |x'-x''|^{1/N}, x',x''\in [0,1], 1 \leq i \leq s,
 \end{equation}
-where the constants $H_i$ are determined by the relation $H_i = 2L_i \sqrt{N+1}$, ${1 \leq i \leq s}$, $L_i$ are the Lipschitz constants from (\ref{eq:4}), and $N$ is the dimensionality of the optimization problem (\ref{eq:1}).
+where the constants $H_i$ are determined by the relation $H_i = 2L_i \sqrt{N+3}$, ${1 \leq i \leq s}$, $L_i$ are the Lipschitz constants from (\ref{eq:4}), and $N$ is the dimensionality of the optimization problem (\ref{eq:1}).
 
@@
 \begin{equation}\label{eq:13}
-A_k (j)=\{ (x_i, y(x_i), z_i=f_j (y(x_i) ): 1 \leq i \leq k,1 \leq j \leq s \},
+A_k (j)=\{ (x_i, y(x_i), z_i=f_j (y(x_i) ): 1 \leq i \leq k \}, 1 \leq j \leq s,
 \end{equation}
@@
 \begin{equation}\label{eq:19}
-m=r M, M=\frac{\max|z_i-z_{i-1}|}{\rho_i}, \rho_i=(x_i-x_{i-1})^{1/N}  ,1 < i \leq k
+m=r M, M=\max_{1 < i \leq k} \frac{|z_i-z_{i-1}|}{\rho_i}, \rho_i=(x_i-x_{i-1})^{1/N}  ,1 < i \leq k
 \end{equation}
@@
 \ForEach {$f(y)\subset \mathcal{F}(y)$ \textbf{and not} $stop$ }
+\State Recalculate $A_k$ to the values of the function $f(y)$;
 \While{$k < kmax$ \textbf{and not} $stop$} 
@@
 	// q is the number of available cores/processors
-	  \State Find the interval with the maximum characteristic 
-		\begin{equation*}
-			t= \argmax_{1<i \leq k} R(i), i \notin T;
-	  \end{equation*}
+	  \State Find the interval with the maximum characteristic $t= \argmax_{1<i \leq k} R(i), i \notin T$;
 		\State Calculate the iteration point $x^{k+i}$ in accordance with (\ref{eq:21});
@@
   \EndIf
+	\EndWhile
 	\State Set $Y = \{ \emptyset \}, Z = \{ \emptyset \}$;
@@
 	\EndFor
-	\EndWhile
+
 \State \textbf{return} $Y$, $Z$, $k$;
@@
 
-When using computational devices with shared memory, the number of cores / processors is quite limited -- the number of processors on one computational node is usually 2, the number of computational cores in one processor usually does not exceed 10-20.
-For increasing possible parallelism, it becomes inevitable to use computational nodes where memory is distributed.
+When using computational devices with shared memory, the number of cores / processors is quite limited -- the number of cores on one computational node usually does not exceed 12-24.
+For increasing possible parallelism, it becomes inevitable to use a distributed memory computer with several computational nodes.
 
@@
 
-In the framework of the proposed approach, a parallel multiple search algorithm (the PMGAS-3 algorithm) is proposed for distributed memory computational systems, the main rules of which are as follows.
+In the framework of the proposed approach, a parallel global optimization algorithm (the PMGAS-3 algorithm) is proposed for distributed memory computational systems, the main rules of which are as follows.
 
-1) Solving the problems of the set $F$ from (\ref{eq:3}) is carried out sequentially in groups; the number of problems in each group coincides with the number of available computational nodes.
+1) Solving the problems of the set $F(y)$ from (\ref{eq:3}) is carried out sequentially in groups; the number of problems in each group coincides with the number of available computational cores $p*q$.
 
-2) Each problem of the current group is solved on its own computational node using the PMGSA-1 algorithm.
+2) Each computational node is applied for solving q optimization problems using the PMGSA-2 algorithm.
 
-At the end of each parallel iteration of the global search, the results of the iteration $(x^{k+j}, z^{k+j})$, $1 \leq j \leq q$, of each processor are transferred to all available nodes of the computational system with the necessary transformation of the values of minimized functions in accordance with the transformation $Z$ from (\ref{eq:5}).
+1) At the end of each parallel iteration of the global search, the results of the iteration $(x^{k+j},z^{k+j} )$, $1 \leq j \leq q$, of each node are transferred to all available computational nodes with the necessary recalculation of the values of functions being minimized in accordance with the transformation $Z$ from (\ref{eq:5}).
 
-Before performing a new iteration, each computational node accepts the search data transmitted to it and adds this data to the search information $A_k$ from (\ref{eq:13}).
+2) Before performing a new iteration, each computational node accepts the search data transmitted to it and adds this data to the search information $A_k (is)$, $1\leq is \leq q$, from (\ref{eq:13}).
 
-In this way, in keeping with the above rules, the PMGSA-3 algorithm is distributed -- the problems of the set $F$ are solved independently on different computational nodes.
-During the global search process, computational nodes exchange the calculation results that are obtained; after each iteration of the global search, the search information for each problem being solved is updated with $q$ calculated values of the minimized function.
-Thus, the more computational nodes there are (and accordingly, simultaneously solved problems), the faster as a rule, each optimization problem is solved.
-This is evidenced, in particular, by the results of numerical experiments that were performed (see Section \ref{eq:5}).
-An example how PMGSA-3 performs an optimization iteration is given in Fig.
+In this way, in keeping with the above rules, the PMGSA-3 algorithm is distributed -- the problems of the set $F(y)$ are solved independently on different computational nodes.
+During the global search process, computational nodes exchange the calculation results that are obtained; after each iteration of the global search, the search information for each problem being solved is updated with $q$ calculated values of the function being minimized.
+Thus, the more computational nodes there are (and accordingly, the problems that are solved simultaneously), the faster, as a rule, each optimization problem is solved.
+This is evidenced, in particular, by the results of numerical experiments that were performed (see Section \ref{sec:5}).
+An example of how PMGSA-3 performs an optimization iteration is given in Fig.
 \ref{fig:4}.
@@
 
-Numerical experiments were performed on the supercomputers Lobachevsky (University of Nizhni Novgorod), Lomonosov (Moscow State University), MVS-10P (Joint Supercomputer Center of RAS) and supercomputers Endeavor.
-The numerical results were obtained by using the following computational nodes: 2 Intel Xeon Platinum 8260L, 2.4 GHz, 256 GB RAM (i.e.
+Numerical experiments were performed on the supercomputers Lobachevsky (University of Nizhni Novgorod) and Endeavour (Intel Cluster).
+The first supercomputer was used for software development.
+The numerical results were obtained by the second one with the following computational nodes: 2 Intel Xeon Platinum 8260L, 2.4 GHz, 256 GB RAM (i.e.
 a total of 48 CPU cores were available on each node).
 The executable program code was built by using the Intel Parallel Studio XE 2019 software package.
-The numerical experiments were performed using the Globalizer system \cite{c40}.
+The numerical experiments were performed using the Globalizer system \cite{c36}.
+This system offers an opportunity to solve computationally intensive multicriteria multiextremal optimization problems using the parallel global search algorithms presented in the paper.
 
 First, we present the results achieved when comparing the MGSA algorithm with several multicriteria optimization algorithms \cite{c37}.
-In carrying out the experiments, the two-criteria test problem proposed in \cite{c36} was used:
+In carrying out the experiments, the test two-criteria problem proposed in \cite{c38} was used:
 
-\begin{equation}\label{eq:32}
+\begin{equation}\label{eq:26}
 f_1(y)=(y_1-1) y_2^2+1,f_2 (y)=y_2, 0 \leq y_1, y_2 \leq 1.
@@
 
+This problem can be considered as an example of MGO, since, to approximate the Pareto set, the MGSA algorithm was applied for solving 50 subproblems $G (\lambda, y)$ from (\ref{eq:7}) for different values of convolution coefficients $\lambda$, uniformly distributed in the interval [0,1].
+
 Solving the MCO problem is understood to mean the construction of a numerical approximation of the Pareto domain.
-To assess the quality of the approximation, the completeness and uniformity of coverage of the Pareto domain were compared using the following two indicators \cite{c15,c36}:
+To assess the quality of the approximation, the completeness and uniformity of coverage of the Pareto domain were compared using the following two indicators  \cite{c28,c37}:
 \begin{itemize}
@@
 
-During the experiments, five multicriteria optimization algorithms were compared: the Monte-Carlo (MC) method, the genetic algorithm SEMO from the PISA library \cite{c37,c38}, the Non-uniform coverage (NUC) method \cite{c36}, the bi-objective Lipschitz optimization (BLO) method \cite{c37} and the MGSA algorithm.
-For the MGSA  algorithm, 50 subproblems $G(\lambda, y)$ from (\ref{eq:32}) were solved for different values of convolution coefficients $\lambda$, uniformly distributed in the interval $[0,1]$.
-The results of the experiments are presented in Table \ref{tab:1}.
+During the experiments, five multicriteria optimization algorithms were compared: the Monte-Carlo (MC) method, the genetic algorithm SEMO from the PISA library \cite{c39}, the Non-uniform coverage (NUC) method \cite{c37}, the bi-objective Lipschitz optimization (BLO) method \cite{c38} and the MGSA algorithm.
+The results of solving the problem (\ref{eq:26}) for all these methods (except for MAGCS) were obtained in \cite{c38} -- see Table \ref{tab:1}.
 
@@
 
-These numerical results show that the MGSA algorithm has a noticeable advantage in almost all performance indicators when compared with the methods of multicriteria optimization being considered, even when solving relatively simple MCO problems.
+These numerical results show that the MGSA algorithm has a noticeable advantage in almost all performance indicators when compared with the methods of multicriteria optimization being considered, even when solving relatively simple MGO problems.
 
@@
 In the experiments, the MGO problems were generated as problems seeking a set of efficient solutions to the MCO problems using the minimax convolution $G (\lambda, y)$ from (\ref{eq:7}) for different values of the convolution coefficients $\lambda$, uniformly distributed in the domain $\Lambda$.
-Given the initial assumption about the high complexity of computing the values of minimized functions of the set $F$ from (\ref{eq:3}), the efficiency of the proposed methods is evaluated by the number of optimization iterations performed by the methods to achieve the required accuracy under the stopping condition from (\ref{eq:22}).
+Given the original assumption about the high complexity of computing the values of functions being minimized of the set $F$ from (\ref{eq:3}), the efficiency of the proposed methods is evaluated by the number of optimization iterations performed by the methods to achieve the required accuracy under the stopping condition from (\ref{eq:22}).
 
@@
 $N = 4$, $m = 5$.
-In the experiments, 100 MCO problems were solved, in which multiextremal functions obtained by using the GKLS generator \cite{c39} were used as the efficiency criteria; for each problem 100 efficient solutions were calculated (i.e., a total of $10,000$ subproblems $G (\lambda, y)$ were solved).
+In the experiments, 100 MCO problems were solved, in which multiextremal functions obtained by using the GKLS generator \cite{c40} were used as the efficiency criteria; for each problem 100 efficient solutions were calculated (i.e., a total of 10,000 subproblems $G (\lambda, y)$ were solved).
 The results obtained (the number of search iterations performed before the stopping condition was fulfilled) were averaged out over the number of MCO problems solved.
-In solving the problems, the search accuracy $\varepsilon = 0.05$ from (\ref{eq:22}) and the reliability of the method $r = 5.6$ from (\ref{eq:19}) were used.
+In solving the problems, the search accuracy $\varepsilon = 0.05$ from (\ref{eq:22}) and the reliability parameter $r = 5.6$ from (\ref{eq:19}) were used.
 
 The results of the numerical experiments are presented in Table \ref{tab:2}.
-The first three rows of the table contain information about the computational resources used during the experiments.
-The first row marked as ``$p$'' indicates the number of computational nodes used.
-The number of cores used within a single node is presented in the second row marked as ``$q$''.
-The total number of cores used is presented in the third row marked as ``$p*q$''.
+The first three rows of the Table contain information about the computational resources used during the experiments.
+The first row marked as ``p'' indicates the number of computational nodes used.
+The number of cores used within a single node is presented in the second row marked as ``q''.
+The total number of cores used is presented in the third row marked as ``p*q''.
 The rows four through six show the average number of iterations performed by PMGSA-1,2,3 for solving a single MCO problem accordingly.
@@
   \includegraphics[width=0.7\linewidth]{fig5}
-  \caption{Speedup and efficiency of the PMGSA-3 algorithm in solving five-criteria four-dimensional MKO problems}
+  \caption{Speedup and efficiency of the PMGSA-3 algorithm in solving five-criteria four-dimensional MCO problems}
   \label{fig:5}
@@
 For solving these problems the parallel PMGSA-3 method was applied.
-When performing the experiments, the accuracy of the method $\varepsilon = 0.1$ from (\ref{eq:22}) and the reliability of the method $r = 5.6$ were used.
+When performing the experiments, the accuracy of the method $\varepsilon = 0.1$ and the reliability of the method $r = 5.6$ were used.
 The results of the experiments are presented in Table \ref{tab:3}.
 
-The first three rows of the table, as before, present the number of computational nodes used, the number of cores per a node and the total number of computational cores.
+The first three rows of the Table, as before, present the number of computational nodes used, the number of cores per a node and the total number of computational cores.
 The last three rows correspond to the number of global search iterations performed, the speedup achieved, and the overall reduction of executed iterations, taking into account the reuse of information (the average number of iterations performed by the sequential GSA algorithm without the reuse of search information for solving one MCO problem is, on average, 3,131,284.8).
+
 Based on the results of the experiments, it can be seen that when the dimensionality of the problem is increased, the efficiency of parallel computations using the PMGSA-3 algorithm also increases (when using 480 cores, the speedup is 472.9).
@@
 \centering
-\caption{Results of experiments to solve the five-criterion six-dimensional MCO problem using the parallel PMGSA-3 method}
+\caption{Results of numerical experiments to solve the five-criterion six-dimensional MCO problem using the parallel PMGSA-3 method}
 \label{tab:3}
@@
 
+In order to demonstrate the efficiency of the proposed approach, a problem of vibration isolation for a system with several degrees of freedom consisting of an isolated base and an elastic body has been solved.
+In this problem it is assumed that the protected object can be represented as a multi-mass mechanical system consisting of several material points connected by vibration damping elements.
+As the criteria, the maximum deformation and maximum displacement of the object relative to the base were minimized \cite{c41}.
+The number of variable parameters is $N=3$.
+
+To construct an approximation of the Pareto set, 50 multiextremal functions $G (\lambda, y)$ from (\ref{eq:7}) were minimized for different values of the convolution coefficients $\lambda$, uniformly distributed in the domain $\Lambda$.
+When performing the experiments, the search accuracy $\varepsilon = 0.025$ and the reliability parameter $r = 3$ were used.
+The results of the experiments are presented in Table \ref{tab:4}.
+
+\begin{table}[]
+\centering
+\caption{Results of numerical experiments to solve the applies problem}
+\label{tab:4}
+\begin{adjustbox}{max width=\textwidth}
+\begin{tabular}{ccccccc}
+\hline
+Computational                     & p                     & 1       & 1      & 1       & 1       & 5       \\
+resources                         & q                     & 1       & 1      & 48      & 48      & 48      \\
+used                              & p*q                   & 1       & 1      & 48      & 48      & 240     \\ \hline
+\multicolumn{2}{c}{Method}                                & GSA     & MGSA   & PMGSA-1 & PMGSA-2 & PMGSA-3 \\
+\multicolumn{2}{c}{Iterations}                            & 703,802 & 13,120 & 403     & 237     & 116     \\
+\multicolumn{2}{c}{HV}                                    & 1.46    & 1.45   & 1.38    & 1.88    & 1.39    \\
+\multicolumn{2}{c}{DU}                                    & 501.03  & 499.43 & 500.46  & 498.73  & 500.39  \\
+\multicolumn{2}{c}{Speedup}                               & -       & 1      & 32.6    & 55.4    & 113.1   \\
+\multicolumn{2}{c}{\begin{tabular}[c]{@{}c@{}}Overall reduction \\ of iteration number\end{tabular}} & 1       & 53.6   & 1,746.4 & 2,969.6 & 6,067.3 \\ \hline
+\end{tabular}
+\end{adjustbox}
+\end{table}
+
+The results of the numerical experiments demonstrate that all methods have found the sufficient approximation of the Pareto domain.
+The lowest number of iterations performed the PMGSA-2 method.
+In Fig.
+\ref{fig:6}, the computed approximation of the Pareto domain obtained by the PMGSA-3method is presented.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.7\linewidth]{fig6}
+  \caption{Approximation of the Pareto domain for the problem of vibration isolation constructed by using PMAGS-2 method}
+  \label{fig:6}
+\end{figure}
+
 \section{Conclusion}\label{sec:6}
 
-The paper considers a new class of optimization problems -- multiple global optimization problems, in which minimized functions can be multiextremal, and calculating function values may require a huge amount of calculations.
-In addition, the set of problems to be solved can change dynamically during calculations by adding or removing solvable global optimization problems.
+The paper considers a new approach to solving multiple global optimization problems, in which minimized functions can be multiextremal, and calculating function values may require a huge amount of calculations.
 Problems of this kind are encountered, for example, when it is necessary to find several efficient solutions of multicriteria optimization problems or when part of the variable parameters of optimization problems can take only a number of discrete values.
@@
 Results of computational experiments show that the proposed approach allows us to significantly -- by tens and hundreds of times -- reduce the computational complexity of solving multiple global optimization problems.
+
 In general, one can observe that the proposed approach is promising and requires further research.

### 2021-03-05 2020/JPDC_MCO/2020_JPDC.tex
1. исправил благодарность

@@
 \section*{Acknowledgements} 
-This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''
+The work was supported by the Ministry of Science and Higher Education of the Russian Federation, agreement No 075-15-2020-808
 

### 2021-03-06 2020/JPDC_MCO/2020_JPDC.tex
JPDC правки

@@
 Unfortunately, the problem of choosing initial points still remains open -- even using a significant number of initial points does not guarantee a reliable solution of global optimization problems.
-Another obvious approach is to construct in the search domain  a number of grids in whose nodes values of functions to be minimized are computed \cite{c2,c3,c4,c5}.
+Another obvious approach is to construct a grid in the search domain, at the nodes of which the values of functions to be minimized are computed \cite{c2,c3,c4,c5}.
 To reduce computational complexity of uniform grids, this approach envisages different methods for construction of non-uniform grids, which would be denser only in the vicinity of the global minimum.
@@
 Such problems are encountered, for example, when one needs to find several efficient solutions of multicriteria optimization problems or when part of the variable parameters in the optimization problem can take only a certain number of discrete values.
-For the first time, the formulation of such multiple global optimization (MGO was considered in \cite{c25,c26}, and then was investigated in \cite{c27,c28}.
+For the first time, the formulation of such multiple global optimization (MGO) was considered in \cite{c25,c26}, and then was investigated in \cite{c27,c28}.
 The relatively limited attention given to MGO problems is most likely due to the high computational complexity involved for solving even one GO problem.
@@
 However, this approach results in a many-fold increase in the computational complexity of solving such problems.
-Meanwhile, MGO problems may have a certain informational interrelation -- a similar connectivity is observed, for example, when methods based on scalarization of the vector efficiency criterion are applied to solve multicriteria optimization (MCO) problems.
+Meanwhile, individual GO problems that are part of the same MGO problem may have a certain informational interrelation -- a similar connectivity is observed, for example, when methods based on scalarization of the vector efficiency criterion are applied to solve multicriteria optimization (MCO) problems.
 In this case, it becomes necessary to solve multiple scalar (in the general case, global) optimization problems in order to find several efficient decisions  for MCO problems -- see, for example, \cite{c29,c30,c31,c32,c33}.
-Such interconnectedness of MGO problems can be used to reduce significantly the amount of computation performed when solving the entire set of problems.
+Such interconnectedness of GO problems can be used to reduce significantly the amount of computation performed when solving the entire set of problems.
 
-The approach proposed in this paper assumes that the computational values of any function of the set of MGO problems can be recalculated to the values of any other functions of the set without any time consuming calculations.
+The approach proposed in this paper assumes that the computational values of any function of the same MGO problem can be recalculated to the values of any other functions of the set without any time consuming calculations.
 In such cases, all the search information obtained in solving a particular problem can be used in solving all other problems in the set.
-Such repeated reuse of search information can significantly reduce the amount of computation performed, up to just a few iterations when solving the next MGO problems.
+Such repeated reuse of search information can significantly reduce the amount of computation performed, up to just a few iterations when solving the next GO problems.
 In addition, the presence of informational connectivity significantly expands the possibilities of a parallel solution for MGO problems.
-The paper proposes two new parallel computational schemes: a cooperative scheme, which means the search information between the problems is exchanged when solving several MGO problems in parallel, and a competitive scheme, which occurs when MGO problems solved in parallel compete with each other for the use of available computational resources.
+The paper proposes two new parallel computational schemes: a cooperative scheme, which means the search information between the problems is exchanged between computational devices with distributed memory used for solving several GO problems in parallel, and a competitive scheme, which occurs when GO problems solved in parallel compete with each other for the use of available computational resources with shared memory.
 
@@
 
-As a result, if the values of efficiency criteria $g_i(y)$, $1 \leq i \leq s$, are computed at some point $y=\bar{y}$ of the search domain $D$, then the value of the scalar criterion $G(\lambda,y)$ can be computed at the point $\bar{y} \in D$ for any values of the coefficients $\lambda \in \Lambda$.
+As a result, if the values of efficiency criteria $g_i(y)$, $1 \leq i \leq m$, are computed at some point $y=\bar{y}$ of the search domain $D$, then the value of the scalar criterion $G(\lambda,y)$ can be computed at the point $\bar{y} \in D$ for any values of the coefficients $\lambda \in \Lambda$.
 In this case the transformation of $Z$ from (\ref{eq:5}) is determined by the expression (\ref{eq:7}) at fixed values of the criteria $g_l (y)$,$1 \leq l \leq m$) \cite{c27,c28}.
@@
 \end{equation}
-where $j$, $1 \leq j \leq s$, is the number of the current function being minimized $f_j (y)$ of the set $F$ from (\ref{eq:3}), k is the number of iterations of the global search, $y^i \in D$, $1 \leq i \leq k$,  are points in which the values of the optimized problems of the set $F$ were computed, and $f^i$, $1 \leq i \leq k$, are the values of the function $f_j (y)$ at these points (the procedure for calculating the value of the function will be called a trial).
+where $j$, $1 \leq j \leq s$, is the number of the current function being minimized $f_j (y)$ of the set $F$ from (\ref{eq:3}), $k$ is the number of iterations of the global search, $y^i \in D$, $1 \leq i \leq k$,  are points in which the values of the optimized problems of the set $F$ were computed, and $f^i$, $1 \leq i \leq k$, are the values of the function $f_j (y)$ at these points (the procedure for calculating the value of the function will be called a trial).
 Global optimization methods can use this search information, thereby ensuring the ability to solve each successive problem of the $F$ set, taking into account the results of all previous calculations.
@@
   \includegraphics[width=\linewidth]{fig1}
-  \caption{Examples of approximations of the Peano space-filling curves (the approximation density is equal to 2, 3 and 4 correspondingly) \cite{c5}}
+  \caption{Examples of approximations of the Peano space-filling curves (the approximation density is equal to 2, 3 and 4 correspondingly) \cite{c6}}
   \label{fig:1}
@@
 \end{equation}
-where $H$ the values $z_i$, $1<i \leq k$, are from $A_k$, H is the H\"older constant from (\ref{eq:12}) for the reduced global optimization problem to be solved and $N$ is the dimensionality of the problem from (\ref{eq:1}).
+where the values $z_i$, $1<i \leq k$, are from $A_k$, $H$ is the H\"older constant from (\ref{eq:12}) for the reduced global optimization problem to be solved and $N$ is the dimensionality of the problem from (\ref{eq:1}).
 For the multidimensional generalized global search algorithm (GSA) \cite{c6,c23}, developed in the framework of the information-statistical approach, the characteristic has the form
@@
 
-A detailed description of the GSA algorithm is given in Algorithm \ref{alg:1}(see also \cite{c6}).
+A detailed description of the GSA algorithm is given in Algorithm \ref{alg:1} (see also \cite{c6}).
 The inputs of the algorithm are the function to be minimized $f(y)$\footnote{ }, the search domain $D$ from (\ref{eq:2}), the reliability parameter $r > 1$, the required accuracy $\varepsilon > 0$ and the maximum permissible number of global search iterations $kmax$.
@@
 
-An obvious way to apply parallel computing is to solve the problems of the set F in parallel.
+An obvious way to apply parallel computing is to solve the problems of the set $F$ in parallel.
 Undoubtedly this approach is straightforward and does not require any additional efforts to be  implemented.
-In this case, if a sufficient number of computational devices (cores / processors) is available, the time needed to solve all problems from the F set will be determined by the time used in solving the most time-consuming problem.
+In this case, if a sufficient number of computational devices ({cores / processors}) is available, the time needed to solve all problems from the set $F$ will be determined by the time used in solving the most time-consuming problem.
 However, this time may turn out to be quite long and necessitate the use of parallel computing to solve even one global optimization problem.
@@
 To implement this approach, it is necessary to ensure that at each iteration of the global optimization problem, not one, but several points in the search domain $D$ are selected for parallel calculation of the values of the functions to be minimized at these points.
-In keeping with the fact that characteristically represented global search algorithms use the characteristics of the intervals $R(i)$, $1 < i \leq k$, from (\ref{eq:15}), to assess the possibility of locating the global minimum in the subinterval of the interval [0,1], it is most reasonable to select the points of the next iteration of the global search in the intervals with the maximum values of the characteristics \cite{c6}.
+In keeping with the fact that characteristically represented global search algorithms use the characteristics of the intervals $R(i)$, $1 < i \leq k$, from (\ref{eq:15}), to assess the possibility of locating the global minimum in the subintervals of the interval [0,1], it is most reasonable to select the points of the next iteration of the global search in the intervals with the maximum values of the characteristics \cite{c6}.
 
 A parallel generalization of the MGSA algorithm (PMGSA-1) is given in Algorithm \ref{alg:2}.
+
+Input data of this algorithm are the set of functions $F(y)$ from (\ref{eq:3}), search domain $D$ from (\ref{eq:2}), reliability parameter $r>1$, required accuracy $\varepsilon > 0$, maximum possible number of calculations of the values of the functions to be minimized $kmax$ and the number of available computational devices ({cores/processors}) with shared memory $q>1$.
+The results of executing the algorithm are the sets of estimates of global minimum points of the functions being minimized $Y$ and values of the functions $Z$ in these points, and the total number of executed iterations of global search $k$.
+As before, in the course of calculations we use the search information $A_k$, which contains the points of earlier iterations of global search $x_i$, $1 \leq i \leq k$, ordered in ascending order, and values of the current function being minimized $z_i=f(y(x_i))$, $1 \leq i \leq k$.
 
@@
 Each optimization problem will compete for available computational resources -- the distribution of resources between problems can be carried out, as before, based on the values of interval characteristics.
-To calculate the characteristics, the problems solved in parallel have their own separate sets of search information $A_k (is)$, $1 \leq is \leq ns$, from (\ref{eq:13}), which match the set of iteration points $x_i$, $1 \leq i \leq k$, of the global search, but differ, accordingly, in the values $z_i (is)$, $1 \leq is \leq k$, of the minimized functions $f_is (y)$, $1 \leq is \leq ns$, from the set $F(y)$.
-As a result, in this case, to specify the intervals where the characteristics are computed, two indices $( is, t )$ are required, where $is$, $1 \leq is \leq ns$, is the number of the function to be minimized  $f_is (y)$ and  $t$, $1 \leq t \leq k$, is the number of the interval in the search information $A_k (is)$.
+To calculate the characteristics, the problems solved in parallel have their own separate sets of search information $A_k (is)$, $1 \leq is \leq ns$, from (\ref{eq:13}), which match the set of iteration points $x_i$, $1 \leq i \leq k$, of the global search, but differ, accordingly, in the values $z_i (is)$, $1 \leq i \leq k$, of the minimized functions $f_{is} (y)$, $1 \leq is \leq ns$, from the set $F(y)$.
+As a result, in this case, to specify the intervals where the characteristics are computed, two indices $( is, t )$ are required, where $is$, $1 \leq is \leq ns$, is the number of the function to be minimized  $f_{is}(y)$ and  $t$, $1 \leq t \leq k$, is the number of the interval in the search information $A_k (is)$.
 
-It should also be noted that when implementing this approach to compare the interval characteristics in different global optimization problems, the values of the function being minimized in the available search information should be normalized\footnote{Without loss of generality, we will use sequential numbering of problems solved in parallel from 1 to $ns$.}
+It should also be noted that when implementing this approach to compare the interval characteristics in different global optimization problems, the values of the function being minimized in the available search information should be normalized\footnote{After transformation (\ref{eq:25}), $z_k^*=0$ and $m_k=1$ are carried out for all $ns$ parallel problems.}
 \begin{equation}\label{eq:25}
@@
   \ForEach { $is = 1..ns$ }
-	  \State Calculate the estimate of the H\"older constant $m$ from (\ref{eq:19}) using $A_k$;
+	  \State Calculate the estimate of the H\"older constant $m(is)_k$ from (\ref{eq:19}) using $A_k(is)$;
 		\State Find the estimate of the global minimum value $z(is)_k^*$ of the function $f_{is}(y)$ in accordance with (\ref{eq:23});
@@
 	// q is the number of available cores/processors
-	  \State Find the interval with the maximum characteristic 
+	  \State Find the interval with the maximum characteristic using $A_k' (is)$
 		\begin{equation*}
-			R_{is} (t)= \argmax_{1 \leq is \leq ns, 1 < i \leq k} R(i),i \notin T 
+			R_{is} (t)= \argmax_{1 \leq is \leq ns, 1 < i \leq k} R(i),i \notin T;
 	  \end{equation*}
-		using $A_k' (is)$;
+
 		\State Calculate the iteration point $x^{k+i}$ using $A_k' (is)$ in accordance with (\ref{eq:21});
-    \State Insert $t$ in $T$, is in $IS$ and $x^{k+i}$ in $XK$;
+    \State Insert $t$ in $T$, $is$ in $IS$ and $x^{k+i}$ in $XK$;
 
@@
 Let us briefly explain the rules for performing a parallel iteration of the global search.
-The characteristics of the intervals are calculated for all ns problems to be solved in parallel, and q intervals with the maximum values of the characteristics are selected among them (lines 12-16).
-As a result, the available computational resources are distributed dynamically among problems -- computational devices can be distributed evenly between problems, or they can be allocated to only one problem with the best interval characteristics.
+The characteristics of the intervals are calculated for all $ns$ problems to be solved in parallel, and $q$ intervals with the maximum values of the characteristics are selected among them (lines 12-16).
+As a result, the available computational resources are distributed dynamically among problems -- computational devices can be distributed evenly between different problems, or they can be allocated to only one problem with the best interval characteristics.
 After calculating the values of the functions being minimized, the new search data obtained is added to the search information of all solved problems; regardless of where the intervals for iterating the global search were selected, the search information of all solved problems is updated with a full set of new search data (lines 19-24).
@@
 In any case, getting additional search information from simultaneously solved problems can significantly reduce the time needed to solve each global search problem.
-This conclusion is rather obvious if the global minima points of the functions $f_i (y)$, $1 \leq i \leq s$, of the set $F$ from (\ref{eq:3}) are situated in the same small neighborhood of the search domain $D$ from (\ref{eq:2}).
-But even in the case when the global minima points of different functions differ much from each other, the values of different functions $f_i (y)$, $1 \leq i \leq s$, of the set $F$ computed in the process of global search can be used for each function separately.
-Thus, in PMGSA-2 the execution of one parallel iteration using $q$ computational units adds $q$ computation results.
+This conclusion is rather obvious if the global minima points of the functions $f_i (y)$, $1 \leq is \leq ns$, of the set $F$ from (\ref{eq:3}) are situated in the same small neighborhood of the search domain $D$ from (\ref{eq:2}).
+But even in the case when the global minima points of different functions differ much from each other, the values of different functions $f_i (y)$, $1 \leq is \leq ns$, of the set $F$ computed in the process of global search can be used for each function separately.
+Thus, in PMGSA-2 the execution of one parallel iteration using $q$ computational devices adds $q$ computation results.
 Increasing the amount of available search information allows adaptive global search algorithms to select points for the next iterations of the global search in a more reasonable manner.
-In the limiting case, when performing sufficiently many $iterations \gg 1$, the amount of search information can be sufficient to find the global minimum point of the next problem in the set $F$ without any additional iterations.
+In the limiting case, when performing sufficiently many iterations $k \gg 1$, the amount of search information can be sufficient to find the global minimum point of the next problem in the set $F$ without any additional iterations.
 This effect is proved also by the results of numerical experiments in Section \ref{sec:5}.
@@
 
-When using computational devices with shared memory, the number of cores / processors is quite limited -- the number of cores on one computational node usually does not exceed 12-24.
-For increasing possible parallelism, it becomes inevitable to use a distributed memory computer with several computational nodes.
+When using computational devices with shared memory, the number of {cores/processors} is quite limited -- the number of cores on one computational node usually does not exceed 12-24.
+For increasing possible parallelism, it becomes inevitable to use a distributed memory computational system with several computational nodes.
 
-Let $p>1$ be the number of computational nodes with distributed memory, each of which has $q>1$ computational cores (that is, the total number of cores in a computational system is $p*q$).
+Let $p>1$ be the number of computational devices with distributed memory, each of which has $q>1$ computational cores (that is, the total number of cores in a computational system is $p*q$).
 
@@
 
-2) Each computational node is applied for solving q optimization problems using the PMGSA-2 algorithm.
+2) Each computational node is applied for solving $q$ optimization problems using the PMGSA-2 algorithm.
 
@@
 In this way, in keeping with the above rules, the PMGSA-3 algorithm is distributed -- the problems of the set $F(y)$ are solved independently on different computational nodes.
-During the global search process, computational nodes exchange the calculation results that are obtained; after each iteration of the global search, the search information for each problem being solved is updated with $q$ calculated values of the function being minimized.
+During the global search process, computational nodes exchange the calculation results that are obtained; after each iteration of the global search, the search information for each problem being solved is updated with $p*q$ calculated values of the function being minimized.
 Thus, the more computational nodes there are (and accordingly, the problems that are solved simultaneously), the faster, as a rule, each optimization problem is solved.
@@
 Numerical experiments were performed on the supercomputers Lobachevsky (University of Nizhni Novgorod) and Endeavour (Intel Cluster).
-The first supercomputer was used for software development.
+The first supercomputer was used for optimization software system.
 The numerical results were obtained by the second one with the following computational nodes: 2 Intel Xeon Platinum 8260L, 2.4 GHz, 256 GB RAM (i.e.
@@
 \centering
-\caption{Comparison of the effectiveness of multicriteria optimization algorithms \cite{c37})}
+\caption{Comparison of the effectiveness of multicriteria optimization algorithms)}
 \label{tab:1}
@@
 
-These numerical results show that the MGSA algorithm has a noticeable advantage in almost all performance indicators when compared with the methods of multicriteria optimization being considered, even when solving relatively simple MGO problems.
+These numerical results show that the MGSA algorithm has a noticeable advantage in all performance indicators when compared with the methods of multicriteria optimization being considered, even when solving relatively simple MGO problems.
 
@@
 The first three rows of the Table contain information about the computational resources used during the experiments.
-The first row marked as ``p'' indicates the number of computational nodes used.
-The number of cores used within a single node is presented in the second row marked as ``q''.
-The total number of cores used is presented in the third row marked as ``p*q''.
+The first row marked as ``$p$'' indicates the number of computational nodes used.
+The number of cores used within a single node is presented in the second row marked as ``$q$''.
+The total number of cores used is presented in the third row marked as ``$p*q$''.
 The rows four through six show the average number of iterations performed by PMGSA-1,2,3 for solving a single MCO problem accordingly.
@@
 The results of the numerical experiments demonstrate that all methods have found the sufficient approximation of the Pareto domain.
-The lowest number of iterations performed the PMGSA-2 method.
+The lowest number of iterations performed the PMGSA-3 method.
 In Fig.
-\ref{fig:6}, the computed approximation of the Pareto domain obtained by the PMGSA-3method is presented.
+\ref{fig:6}, the computed approximation of the Pareto domain obtained by the PMGSA-2 method is presented.
 
@@
 The availability of informational connectivity significantly expands the possibilities of parallel solutions of the entire set of multiple global optimization problems.
-The paper proposes two parallel computational schemes: a cooperative scheme, when several problems are solved simultaneously, with the search information that is obtained being exchanged between problems, and a competitive scheme, when problems that are simultaneously being solved compete with each other for using the available computational resources.
+The paper proposes two parallel computational schemes: a cooperative scheme, which means the search information between the problems is exchanged between computational devices with distributed memory used for solving several GO problems in parallel, and a competitive scheme, which occurs when GO problems solved in parallel compete with each other for the use of available computational resources with shared memory..
 

### 2021-03-06 2020/JPDC_MCO/2020_JPDC.tex
1

@@
 
-	\State
+	\State Calculate the stop condition
 	\begin{equation}\label{eq:22}
-     stop = (x_t - x_{t-1})^{1/N} \leq \varepsilon.
+     stop = \begin{cases} true, &(x_t - x_{t-1})^{1/N} \leq \varepsilon,\\
+		                      false,&(x_t - x_{t-1})^{1/N} > \varepsilon.
+							\end{cases}		
   \end{equation}
@@
 Undoubtedly this approach is straightforward and does not require any additional efforts to be  implemented.
-In this case, if a sufficient number of computational devices ({cores / processors}) is available, the time needed to solve all problems from the set $F$ will be determined by the time used in solving the most time-consuming problem.
+In this case, if a sufficient number of computational devices ({cores/processors}) is available, the time needed to solve all problems from the set $F$ will be determined by the time used in solving the most time-consuming problem.
 However, this time may turn out to be quite long and necessitate the use of parallel computing to solve even one global optimization problem.

### 2021-03-07 2020/JPDC_MCO/2020_JPDC.tex
JPDC добавил перевод сносок

@@
 A detailed description of the GSA algorithm is given in Algorithm \ref{alg:1} (see also \cite{c6}).
-The inputs of the algorithm are the function to be minimized $f(y)$\footnote{ }, the search domain $D$ from (\ref{eq:2}), the reliability parameter $r > 1$, the required accuracy $\varepsilon > 0$ and the maximum permissible number of global search iterations $kmax$.
+The inputs of the algorithm are the function to be minimized $f(y)$\footnote{In the context of this paper, $f(y)$ can be regarded as one of the functions of the set $F(y)$ from (\ref{eq:3}).}, the search domain $D$ from (\ref{eq:2}), the reliability parameter $r > 1$, the required accuracy $\varepsilon > 0$ and the maximum permissible number of global search iterations $kmax$.
 The output of the algorithm is an estimate of the global minimum point $y_k^* =y(x_k^*)$, the value of the function to be minimized at this point $z_k^* =f(y_k^*)$  and the number of performed global search iterations $k$.
-In the course of calculations, the search information $A_k$\footnote{ } from (\ref{eq:13}) is used, which contains the points of previously performed global search iterations $x_i$, $1 \leq i \leq k$, ordered in ascending order, and the values of the function to be minimized $z_i=f(y(x_i))$, $1\leq i \leq k$.
+In the course of calculations, the search information $A_k$\footnote{The set $A_k$ should be regarded as the search information $A_k(j)$ from (\ref{eq:13}) obtained by minimizing only one function $f(y)$.} from (\ref{eq:13}) is used, which contains the points of previously performed global search iterations $x_i$, $1 \leq i \leq k$, ordered in ascending order, and the values of the function to be minimized $z_i=f(y(x_i))$, $1\leq i \leq k$.
 

### 2021-03-07 2020/JPDC_MCO/2020_JPDC.tex
правки статьи

@@
 In addition, the presence of informational connectivity significantly expands the possibilities of a parallel solution for MGO problems.
-The paper proposes two new parallel computational schemes: a cooperative scheme, which means the search information between the problems is exchanged between computational devices with distributed memory used for solving several GO problems in parallel, and a competitive scheme, which occurs when GO problems solved in parallel compete with each other for the use of available computational resources with shared memory.
+The paper proposes two new parallel computational schemes: a cooperative scheme, which means the search information is exchanged between computational devices with distributed memory used for solving several GO problems in parallel, and a competitive scheme, which occurs when GO problems solved in parallel compete with each other for the use of available computational resources with shared memory.
 
@@
 
-A key condition of this approach is the assumption that any calculated value of any function $f_i(y)$, $1 \leq i \leq s$, of the set F at an arbitrary point $y \in D$ of the search domain $D$ can be transformed to the value of any other function of the set $F$ at the same point $y \in D$ without any time consuming computations, in a time significantly shorter compared to the computation time values of the functions $f_i(y)$, $1 \leq i \leq s$.
+A key condition of this approach is the assumption that any calculated value of any function $f_i(y)$, $1 \leq i \leq s$, of the set $F$ at an arbitrary point $y \in D$ of the search domain $D$ can be transformed to the value of any other function of the set $F$ at the same point $y \in D$ without any time consuming computations, in a time significantly shorter compared to the computation time values of the functions $f_i(y)$, $1 \leq i \leq s$.
 In other words, there must be a transformation $Z$ such that
@@
 To find efficient (unimprovable) solutions, a widely used approach is based on the reduction (scalarization) of the vector criterion $g(y)$ to the general scalar criterion, the minimization of which will lead to the Pareto optimal solution of the MCO problem \cite{c29,c30,c31,c32,c33}.
-So, when using mini-max convolution, the vector criterion $g(y)$ from (\ref{eq:6}) reduces to the scalar criterion
+So, when using minimax convolution, the vector criterion $g(y)$ from (\ref{eq:6}) reduces to the scalar criterion
 \begin{equation}\label{eq:7}
@@
 where $j$, $1 \leq j \leq s$, is the number of the current function being minimized $f_j (y)$ of the set $F$ from (\ref{eq:3}), $k$ is the number of iterations of the global search, $y^i \in D$, $1 \leq i \leq k$,  are points in which the values of the optimized problems of the set $F$ were computed, and $f^i$, $1 \leq i \leq k$, are the values of the function $f_j (y)$ at these points (the procedure for calculating the value of the function will be called a trial).
-Global optimization methods can use this search information, thereby ensuring the ability to solve each successive problem of the $F$ set, taking into account the results of all previous calculations.
+Global optimization methods can use this search information, thereby ensuring the ability to solve each successive problem of the set $F$, taking into account the results of all previous calculations.
 Numerical experiments that were performed (see Section \ref{sec:5}), confirm that such repeated use of search information reduces the number of iterations of the global search required to solve the next problem of the set $F$ (even to the point of performing just a few search iterations).
@@
 In accordance with this theory, first of all, dimensionality reduction is achieved using the Peano space-filling curves (evolvents) $y(x)$, uniquely and continuously mapping the interval [0,1] to the $N$-dimensional domain $D$ -- see, for example, \cite{c6,c23}.
-As a result of this reduction, multidimensional global optimization problems (\ref{eq:1})-(\ref{eq:2}) can be reduced to one-dimensional ones:
+As a result of this reduction, multidimensional global optimization problems (\ref{eq:1})-(\ref{eq:3}) can be reduced to one-dimensional ones:
 
@@
 \end{equation}
-or a more efficient execution of global search algorithms.
+for a more efficient execution of global search algorithms.
 It should be noted that in optimizing the reduced one-dimensional functions $f_i (y(x))$, $1 \leq i \leq s$, many one-dimensional global search algorithms can be used (possibly after some generalization).
@@
   \includegraphics[width=\linewidth]{fig1}
-  \caption{Examples of approximations of the Peano space-filling curves (the approximation density is equal to 2, 3 and 4 correspondingly) \cite{c6}}
+  \caption{Examples of approximations of the Peano space-filling curves (the approximation density is equal to 2, 3 and 4 correspondingly)}
   \label{fig:1}
@@
 
-	\State Calculate the stop condition
+	\State Check the stopping condition
 	\begin{equation}\label{eq:22}
@@
 Moreover, due to the informational connectivity (\ref{eq:5}) of the problems of the set $F$, the search information $A_k$ can be quickly transformed to the values of the next function to be minimized $f(y)$ of the set $F$ and GSA can minimize each successive function $f(y)$, taking into account the results of all previous calculations.
-The GSA algorithm for solving problems of the set $F$ and using search information $A_k$, will be referred to as the multiple global search algorithm (MGSA).
+The GSA algorithm for solving GO problems of the set $F$ and using search information $A_k$, will be referred to as the \textit{Multiple Global Search Algorithm} (MGSA).
 
@@
 This approach is general -- it can be applied to solving any global optimization problems in a parallel mode.
-And such an approach is efficient, since the most time consuming part of the problem is parallelized due to the initial assumption that calculating the values of functions to be minimized is computationally expensive.
+And such an approach is efficient, since the most time consuming part of computations is parallelized due to the initial assumption that calculating the values of functions to be minimized is computationally expensive.
 
@@
 
-Input data of this algorithm are the set of functions $F(y)$ from (\ref{eq:3}), search domain $D$ from (\ref{eq:2}), reliability parameter $r>1$, required accuracy $\varepsilon > 0$, maximum possible number of calculations of the values of the functions to be minimized $kmax$ and the number of available computational devices ({cores/processors}) with shared memory $q>1$.
-The results of executing the algorithm are the sets of estimates of global minimum points of the functions being minimized $Y$ and values of the functions $Z$ in these points, and the total number of executed iterations of global search $k$.
+Input data of this algorithm are the set of functions $F(y)$ from (\ref{eq:3}), the search domain $D$ from (\ref{eq:2}), the reliability parameter $r>1$, the required accuracy $\varepsilon > 0$, the maximum possible number of calculations of the values of the functions to be minimized $kmax$ and the number of available computational devices ({cores/processors}) with shared memory $q>1$.
+The results of executing the algorithm are the sets of estimates of global minimum points of the functions being minimized $Y$ and values of the functions $Z$ at these points, and the total number of executed iterations of global search $k$.
 As before, in the course of calculations we use the search information $A_k$, which contains the points of earlier iterations of global search $x_i$, $1 \leq i \leq k$, ordered in ascending order, and values of the current function being minimized $z_i=f(y(x_i))$, $1 \leq i \leq k$.
@@
 \end{equation}
-where $z(is)_k$, $1 \leq is \leq ns$, are the estimates (\ref{eq:23}) of the minimum values of the functions being minimized, and $m(is)_k$, $1 \leq is \leq ns$, are the estimates (\ref{eq:19}) of the H\"older constants $H_{is}$, $1 \leq is \leq ns$, from (\ref{eq:12}).
+where $z(is)_k^*$, $1 \leq is \leq ns$, are the estimates (\ref{eq:23}) of the minimum values of the functions being minimized, and $m(is)_k$, $1 \leq is \leq ns$, are the estimates (\ref{eq:19}) of the H\"older constants $H_{is}$, $1 \leq is \leq ns$, from (\ref{eq:12}).
 Further, when performing the steps of the algorithm, the converted values $z(is)_i'$, $1 \leq i \leq k$, $1 \leq is \leq ns$, should be used when calculating the characteristics of intervals and points of new iterations of the global search.
@@
   \includegraphics[width=0.7\linewidth]{fig3}
-  \caption{An example of intervals selection by the PMGSA-2 algorithm (Step 1)}
+  \caption{An example of intervals selection by the PMGSA-2 algorithm}
   \label{fig:3}
@@
 
-1) At the end of each parallel iteration of the global search, the results of the iteration $(x^{k+j},z^{k+j} )$, $1 \leq j \leq q$, of each node are transferred to all available computational nodes with the necessary recalculation of the values of functions being minimized in accordance with the transformation $Z$ from (\ref{eq:5}).
+3) At the end of each parallel iteration of the global search, the results of the iteration $(x^{k+j},z^{k+j} )$, $1 \leq j \leq q$, of each node are transferred to all available computational nodes with the necessary recalculation of the values of functions being minimized in accordance with the transformation $Z$ from (\ref{eq:5}).
 
-2) Before performing a new iteration, each computational node accepts the search data transmitted to it and adds this data to the search information $A_k (is)$, $1\leq is \leq q$, from (\ref{eq:13}).
+4) Before performing a new iteration, each computational node accepts the search data transmitted to it and adds this data to the search information $A_k (is)$, $1\leq is \leq q$, from (\ref{eq:13}).
 
@@
 Numerical experiments were performed on the supercomputers Lobachevsky (University of Nizhni Novgorod) and Endeavour (Intel Cluster).
-The first supercomputer was used for optimization software system.
+The first supercomputer was used for optimization software system development.
 The numerical results were obtained by the second one with the following computational nodes: 2 Intel Xeon Platinum 8260L, 2.4 GHz, 256 GB RAM (i.e.
@@
 \centering
-\caption{Comparison of the effectiveness of multicriteria optimization algorithms)}
+\caption{Comparison of the effectiveness of multicriteria optimization algorithms}
 \label{tab:1}
@@
 The total number of cores used is presented in the third row marked as ``$p*q$''.
-The rows four through six show the average number of iterations performed by PMGSA-1,2,3 for solving a single MCO problem accordingly.
+The rows four through six show the average number of iterations performed by PMGSA-1,2,3 for solving a single MCO problem (i.e., 100 GO problems) accordingly.
 
@@
 For solving these problems the parallel PMGSA-3 method was applied.
-When performing the experiments, the accuracy of the method $\varepsilon = 0.1$ and the reliability of the method $r = 5.6$ were used.
+When performing the experiments, the accuracy $\varepsilon = 0.1$ and the reliability $r = 5.6$ were used.
 The results of the experiments are presented in Table \ref{tab:3}.
@@
 \centering
-\caption{Results of numerical experiments to solve the applies problem}
+\caption{Results of numerical experiments to solve the applied problem}
 \label{tab:4}
@@
 The availability of informational connectivity significantly expands the possibilities of parallel solutions of the entire set of multiple global optimization problems.
-The paper proposes two parallel computational schemes: a cooperative scheme, which means the search information between the problems is exchanged between computational devices with distributed memory used for solving several GO problems in parallel, and a competitive scheme, which occurs when GO problems solved in parallel compete with each other for the use of available computational resources with shared memory..
+The paper proposes two parallel computational schemes: a cooperative scheme, which means the search information between the problems is exchanged between computational devices with distributed memory used for solving several GO problems in parallel, and a competitive scheme, which occurs when GO problems solved in parallel compete with each other for the use of available computational resources with shared memory.
 

### 2021-03-08 2020/JPDC_MCO/2020_JPDC.tex
1

@@
 Moreover, due to the informational connectivity (\ref{eq:5}) of the problems of the set $F$, the search information $A_k$ can be quickly transformed to the values of the next function to be minimized $f(y)$ of the set $F$ and GSA can minimize each successive function $f(y)$, taking into account the results of all previous calculations.
-The GSA algorithm for solving GO problems of the set $F$ and using search information $A_k$, will be referred to as the \textit{Multiple Global Search Algorithm} (MGSA).
+The GSA algorithm for solving GO problems of the set $F$ and using search information $A_k$, will be referred to as the Multiple Global Search Algorithm (MGSA).
 
@@
 
-\biboptions{sort&compress}
-\bibliography{mybibfile}
+\begin{thebibliography}{10}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\expandafter\ifx\csname href\endcsname\relax
+  \def\href#1#2{#2} \def\path#1{#1}\fi
+
+\bibitem{c1}
+R.~Arora, Optimization: Algorithms and Applications, CRC Press, 2015.
+
+\bibitem{c2}
+M.~Locatelli, F.~Schoen, Global Optimization: Theory, Algorithms, and
+  Applications, SIAM, 2013.
+
+\bibitem{c3}
+R.~Paulavi{\v c}ius, J.~{\v Z}ilinskas, Simplicial Global Optimization,
+  Springer Briefs in Optimization, Springer-Verlag, New-York, 2014.
+
+\bibitem{c4}
+X.-S. Yang, Nature-inspired metaheuristic algorithms, Luniver Press, Frome,
+  2008.
+
+\bibitem{c5}
+A.~Zhigljavsky, A.~{\v Z}ilinskas, Stochastic Global Optimization, Springer,
+  2008.
+
+\bibitem{c6}
+R.~Strongin, Y.~Sergeyev, Global optimization with non-convex constraints.
+  Sequential and parallel algorithms, Kluwer Academic Publishers, Dordrecht,
+  2000 (2nd ed. 2013, 3rd ed. 2014).
+
+\bibitem{c7}
+Y.~Sergeyev, D.~Kvasov, Deterministic Global Optimization: An Introduction to
+  the Diagonal Approach, Springer Briefs in Optimization, Springer-Verlag,
+  New-York, 2017.
+
+\bibitem{c8}
+L.~M. Rios, N.~V. Sahinidis, Derivative-free optimization: a review of
+  algorithms and comparison of software implementations, Journal of Global
+  Optimization 56 (2013) 1247--1293.
+\newblock \href {https://doi.org/10.1007/s10898-012-9951-y}
+  {\path{doi:10.1007/s10898-012-9951-y}}.
+
+\bibitem{c9}
+C.~Floudas, C.~Gounaris, A review of recent advances in global optimization,
+  Journal of Global Optimization 45~(3) (2009).
+\newblock \href {https://doi.org/10.1007/s10898-008-9332-8}
+  {\path{doi:10.1007/s10898-008-9332-8}}.
+
+\bibitem{c10}
+R.~Mart{\'i}, Multi-Start Methods, Springer, US, Boston, MA, 2003.
+\newblock \href {https://doi.org/10.1007/0-306-48056-5_12}
+  {\path{doi:10.1007/0-306-48056-5_12}}.
+
+\bibitem{c11}
+T.~Krityakierne, C.~Shoemaker, {SOMS}: {SurrOgate} {MultiStart} algorithm for
+  use with nonlinear programming for global optimization, International
+  transactions in operational research 24~(5) (2017) 1139--1172.
+
+\bibitem{c12}
+A.~{\v{Z}}ilinskas, J.~Gillard, M.~Scammell, A.~Zhigljavsky, Multistart with
+  early termination of descents, Journal of Global Optimization 79~(2) (2021)
+  447--462.
+\newblock \href {https://doi.org/10.1007/s10898-019-00814-w}
+  {\path{doi:10.1007/s10898-019-00814-w}}.
+
+\bibitem{c13}
+Y.~Gao, W.~Du, G.~Yan, Selectively-informed particle swarm optimization,
+  Scientific Reports 5~(1) (2015) 92--95.
+\newblock \href {https://doi.org/10.1038/srep09295}
+  {\path{doi:10.1038/srep09295}}.
+
+\bibitem{c14}
+A.~Eiben, J.~Smith, Introduction to evolutionary computing, Springer, Berlin,
+  Heidelberg, 2015.
+
+\bibitem{c15}
+M.~Gendreau, Handbook of metaheuristics, Springer, New-York, 2010.
+
+\bibitem{c16}
+H.~Tuy, Convex Analysis and Global Optimization, Springer, 2016.
+
+\bibitem{c17}
+H.~A. Le~Thi, T.~Pham Dinh, {DC} programming and {DCA}: thirty years of
+  developments, Mathematical Programming 169~(1) (2018) 5--68.
+\newblock \href {https://doi.org/10.1007/s10107-018-1235-y}
+  {\path{doi:10.1007/s10107-018-1235-y}}.
+
+\bibitem{c18}
+E.~Hansen, G.~Walster, Global optimization using interval analysis: revised and
+  expanded, CRC Press, Boca Raton, 2003.
+
+\bibitem{c19}
+Y.~Evtushenko, M.~Posypkin, A deterministic approach to global box-constrained
+  optimization, Optimization Letters 7~(4) (2013) 819--829.
+\newblock \href {https://doi.org/10.1007/s11590-012-0452-1}
+  {\path{doi:10.1007/s11590-012-0452-1}}.
+
+\bibitem{c20}
+R.~Paulavi{\v{c}}ius, J.~{\v{Z}}ilinskas, A.~Grothey, Parallel branch and bound
+  for global optimization with combination of {Lipschitz} bounds, Optimization
+  Methods and Software 26~(3) (2011) 487--498.
+\newblock \href {https://doi.org/10.1080/10556788.2010.551537}
+  {\path{doi:10.1080/10556788.2010.551537}}.
+
+\bibitem{c21}
+A.~Gimbutas, A.~{\v{Z}}ilinskas, An algorithm of simplicial {Lipschitz}
+  optimization with the bi-criteria selection of simplices for the bi-section,
+  Journal of Global Optimization 71~(1) (2018) 115--127.
+\newblock \href {https://doi.org/10.1007/s10898-017-0550-9}
+  {\path{doi:10.1007/s10898-017-0550-9}}.
+
+\bibitem{c22}
+D.~R. Jones, J.~R. R.~A. Martins, The {DIRECT} algorithm: 25 years later,
+  Journal of Global Optimization 79~(3) (2021) 521--566.
+\newblock \href {https://doi.org/10.1007/s10898-020-00952-6}
+  {\path{doi:10.1007/s10898-020-00952-6}}.
+
+\bibitem{c23}
+Y.~Sergeyev, R.~Strongin, D.~Lera, Introduction to global optimization
+  exploiting space-filling curves, Springer Briefs in Optimization, Springer,
+  New-York, 2013.
+
+\bibitem{c25}
+R.~Rockafellar, Lipschitzian properties of multifunctions, Nonlinear Analysis:
+  Theory, Methods {\&} Applications 9~(8) (1985) 867--885.
+
+\bibitem{c26}
+D.~S. Kim, G.~M. Lee, P.~H. Sach, Hartley proper efficiency in multifunction
+  optimization, Journal of Optimization Theory and Applications 120~(1) (2004)
+  129--145.
+\newblock \href {https://doi.org/10.1023/B:JOTA.0000012736.02360.58}
+  {\path{doi:10.1023/B:JOTA.0000012736.02360.58}}.
+
+\bibitem{c27}
+V.~Gergel, E.~Kozinov, Accelerating parallel multicriterial optimization
+  methods based on intensive using of search information, Procedia Computer
+  Science 108 (2017) 1463--1472, {International} Conference on Computational
+  Science, {ICCS} 2017, 12-14 June 2017, Zurich, Switzerland.
+\newblock \href {https://doi.org/10.1016/j.procs.2017.05.051}
+  {\path{doi:10.1016/j.procs.2017.05.051}}.
+
+\bibitem{c28}
+V.~Gergel, E.~Kozinov, Efficient multicriterial optimization based on intensive
+  reuse of search information, J. Glob. Optim. 71~(1) (2018) 73--90.
+\newblock \href {https://doi.org/10.1007/s10898-018-0624-3}
+  {\path{doi:10.1007/s10898-018-0624-3}}.
+
+\bibitem{c29}
+P.~Pardalos, A.~Zilinskas, J.~{\v{Z}}ilinskas, Non-Convex Multi-Objective
+  Optimization, Springer Optimization and Its Applications, Springer, 2017.
+\newblock \href {https://doi.org/10.1007/978-3-319-61007-8}
+  {\path{doi:10.1007/978-3-319-61007-8}}.
+
+\bibitem{c30}
+Y.~Collette, P.~Siarry, Multiobjective Optimization: Principles and Case
+  Studies (Decision Engineering), Springer-Verlag, Berlin, Heidelberg, 2011.
+
+\bibitem{c31}
+E.~Zavadskas, Z.~Turskis, S.~Kildiene, State of art surveys of overviews on
+  {MCDM}/{MADM} methods, Technological and Economic Development of Economy 20
+  (2014) 165--179.
+\newblock \href {https://doi.org/10.3846/20294913.2014.892037}
+  {\path{doi:10.3846/20294913.2014.892037}}.
+
+\bibitem{c32}
+C.~Hillermeier, J.~Jahn, Multiobjective optimization: Survey of methods and
+  industrial applications, Surveys on Mathematics for Industry 11 (2005) 1--42.
+
+\bibitem{c33}
+G.~Eichfelder, Scalarizations for adaptively solving multi-objective
+  optimization problems, Comput. Optim. Appl. 44 (2009) 249--273.
+\newblock \href {https://doi.org/10.1007/s10589-007-9155-4}
+  {\path{doi:10.1007/s10589-007-9155-4}}.
+
+\bibitem{c34}
+S.~Piyavskii, An algorithm for finding the absolute extremum of a function,
+  Comput. Maths. Math. Phys. 12~(4) (1972) 57--67.
+
+\bibitem{c35}
+B.~Shubert, A sequential method seeking the global maximum of a function,
+  {SIAM} Journal on Numerical Analysis 9~(3) (1972) 379--388.
+
+\bibitem{c36}
+V.~Gergel, K.~Barkalov, A.~Sysoyev, Globalizer: A novel supercomputer software
+  system for solving time-consuming global optimization problems, Numerical
+  Algebra, Control {\&} Optimization 8 (2018) 47--62.
+\newblock \href {https://doi.org/10.3934/naco.2018003}
+  {\path{doi:10.3934/naco.2018003}}.
+
+\bibitem{c37}
+L.~Breiman, A.~Cutler, A deterministic algorithm for global optimization,
+  Mathematical Programming 58 (1993) 179--199.
+\newblock \href {https://doi.org/10.1007/BF01581266}
+  {\path{doi:10.1007/BF01581266}}.
+
+\bibitem{c38}
+A.~{\v Z}ilinskas, J.~{\v Z}ilinskas, Adaptation of a one-step worst-case
+  optimal univariate algorithm of bi-objective {Lipschitz} optimization to
+  multidimensional problems, Commun Nonlinear Sci Numer Simulat 21~(1-3) (2015)
+  89--98.
+\newblock \href {https://doi.org/10.1016/j.cnsns.2014.08.025}
+  {\path{doi:10.1016/j.cnsns.2014.08.025}}.
+
+\bibitem{c39}
+S.~Bleuler, M.~Laumanns, L.~Thiele, E.~Zitzler, {PISA-A} platform and
+  programming language independent interface for search algorithms,
+  Evolutionary Multi-Criterion Optimization, {LNCS} 2632 (2003) 494--508.
+\newblock \href {https://doi.org/10.1007/3-540-36970-8_35}
+  {\path{doi:10.1007/3-540-36970-8_35}}.
+
+\bibitem{c40}
+M.~Gaviano, D.~Kvasov, D.~Lera, Y.~Sergeyev, Software for generation of classes
+  of test functions with known local and global minima for global optimization,
+  {ACM} Transactions on Mathematical Software 29~(4) (2003) 469--480.
+
+\bibitem{c41}
+D.~Balandin, M.~Kogan, Multi-objective generalized {H2} control for optimal
+  protection from vibration, {UKACC} 12th International Conference on Control
+  ({CONTROL}), Sheffield (2018) 205--210. \href
+  {https://doi.org/10.1109/CONTROL.2018.8516721}
+  {\path{doi:10.1109/CONTROL.2018.8516721}}.
+
+\end{thebibliography}
 

### 2021-03-16 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
Some answers for reviewers comments

@@
 The PMGAL algorithm proposed in this work (in its sequential version) was compared with the genetic algorithm for solving mixed-integer optimization problems implemented in the Matlab Global Optimization Toolbox.
-Table \ref{tab:1} shows the number of trials it took for these methods to solve a number of mixed-integer test problems taken from \cite{Deep,Floudas}.
+Table \ref{tab:1} shows the number of trials and time it took for these methods to solve a number of mixed-integer test problems taken from \cite{Deep,Floudas}.
+%Russian
+Использованные тестовые задачи содержали от 2 до 4 непрерывных и от 2 до 4 дискретных параметров.
+
 The same accuracy of search $\epsilon = 10^{-2}$ was used for both methods.
+
 All computational experiments were carried out on a computer with an Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM.
@@
 		\cline{2-3} \cline{4-5} 
-		 & $k$ & $t$ &  $k$ & $t$  \\
+		 & Number of trials & Time, sec. &  Number of trials & Time, sec.  \\
 		\hline 
@@
 
-In total, we generated and solved 100 9-dimensional mixed-integer problems of this type.
+In total, we generated and solved 100 mixed-integer problems of this type with 5 discrete and 4 continuous parameters.
 In order to simulate the computational complexity inherent in applied optimization problems, the calculation of the function in all the experiments was complicated by additional calculations that did not change the form of the function and the location of its minima.
@@
 
-The last series of experiments was performed to evaluate the acceleration of the parallel version of the PMGAL algorithm. 
+The last experiment was performed to evaluate the acceleration of the parallel version of the PMGAL algorithm. 
+%Russian
+Параллельный алгоритм был применен для решения той же серии из 100 problems with 5 discrete and 4 continuous parameters.
+
 Table \ref{tab:3} shows the average solution time  $T_{av}$ (in seconds) depending on the involved number of processes $N_{p}$ and threads $N_{t}$ in each process.
@@
 These problems can also include non-convex constraints, which are accounted for using a special index scheme.
+
+%Russian
+Основным предположением, которое обеспечивает применимость предложенного алгоритма, является выполнение условия Липшица для функций задачи при варьировании непрерывных параметров для каждого фиксированного набора дискретных параметров.
+Задачи оптимизации, в которых функции являются липшицевыми, относятся к классу вычислительно-трудоемких задач.
+Любой численный алгоритм решения такой задачи требует вычисления значений функции в узлах некоторой (равномерной или неравномерной) сетки во всей области допустимых изменений параметров.
+Несмотря на то, что предположение липшицевости позволяет провести оценку оптимума на существенно неравномерной (т.е.
+более экономной по сравнению с равномерной) сетке в области поиска, масштаб решаемых задач здесь ограничен.
+За приемлемое время даже с использованием параллельного алгоритма можно решать задачи, в которых порядка 10 непрерывных и 10 дискретных параметров.
+Решение задач с сотнями переменных может быть основано на схеме вложенной оптимизации в предположении, что зависимость целевой функции от части переменных носит локальный характер \cite{Barkalov2020_1}.
+
 We propose a parallel algorithm for solving the problems of this class.
 The basic principle of organizing parallel computations in the algorithm is based on the simultaneous carrying out of several trials at different points of the search domain.
-
 The parallel global optimization algorithm uses a local tuning technique based on the assumption that the multiextremity of the problem is weak.
 This assumption is true, in particular, at the final phase of the search, when one or more trials have already been performed in the vicinity of the global solution, and the only remaining step is to refine the found estimate of the solution.
-
 In its sequential form, the proposed algorithm remains on par with the similar methods implemented in Matlab Global Optimization Toolbox. 
 In a parallel form, the proposed algorithm demonstrates acceptable scalability up to dozens of parallel processes.
+
+%Russian
+Примером сложной прикладной задачи, к которой планируется применить разработанные алгоритмы, является выбор оптимальных параметров конструкции aircraft hot-air anti-icing system для самолетов российского производства.
+В математической модели anti-icing system наряду с параметрами, изменяющимися непрерывно (например, диаметр трубки, подводящей теплый воздух от двигателя в крыло; диаметр выпускных отверстий на данной трубке; расстояние между отверстиями; и т.п.), есть и дискретные параметры (например, число рядов выпускных отверстий, которое из технических ограничений может быть равно 1, 2 или 3).
+Пример подобной постановки задачи см.
+в \cite{Habashi}.
+Использование предложенного в статье алгоритма позволит не проводить оптимизацию для каждого значения дискретного параметра по отдельности, а заранее отсечь неперспективные варианты в процессе поиска.
 
@@
 \bibitem{Zilinskas2011}
-R.~Paulavi\v{c}ius, J.~\v{Z}ilinskas and A.~Grothey, \textit{Parallel branch and bound for global optimization with combination of Lipschitz bounds}.
-Optimization Methods \& Software \textbf{26}(3), 487--498 (2011).
+R.~Paulavi\v{c}ius, J.~\v{Z}ilinskas and A.~Grothey, \textit{Parallel branch and bound for global optimization with combination of Lipschitz bounds}, Optimization Methods \& Software \textbf{26}(3), 487--498 (2011).
 
@@
 \textbf{29}, 469--480 (2003).
+
+\bibitem{Barkalov2020_1}
+K.~Barkalov, I.~Lebedev, M.~Kocheganova, V.~Gergel, \textit{Combining local and global search in a parallel nested optimization scheme}, Communications in Computer and Information Science, \textbf{1263}, 100-112 (2020).
+
+\bibitem{Habashi}
+M.~P.~C. Pellissier, W.~G. Habashi, A.~Pueyo. \textit{Optimization via FENSAP-ICE of aircraft hot-air anti-icing systems}, Journal of Aircraft \textbf{48}(1), 265--276 (2011).
 

### 2021-03-18 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
change layout of figures and tables
Add some comments in Russian

@@
 
-\begin{figure}
+\begin{figure}[ht]
 \begin{minipage}{0.4\linewidth}
@@
 %Russian
-Использованные тестовые задачи содержали от 2 до 4 непрерывных и от 2 до 4 дискретных параметров.
+Использованные тестовые задачи содержали 2 непрерывных и 1 дискретный параметр, который имел от 3 до 6 значений в зависимости от задачи.
+Также в постановку задач входили от 1 до 4 non-convex constraints. 
+Отметим, что наличие сложных ограничений, которые формируют невыпуклую допустимую область, является препятствием к использованию методов, основанных на идеях локальной оптимизации. 
+Генетический алгоритм из Matlab Global Optimization Toolbox использовался с рекомендованными параметрами (например, population size равнялся 50, максимальное число поколений -- 200, crossover fraction -- 0.8).
+The PMGAL algorithm использовался со стандартными параметрами, обеспечивающими сходимость к решению ( $r_\nu$ from (\ref{R}) равнялось 2.3 для всех $\nu = 1,2,3$, the evolvent construction parameter $M=10$).
 
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \begin{minipage}{0.48\linewidth}
@@
 Computational experiments were carried out on the Lobachevsky supercomputer.
-The node of supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.
+The node of supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs (16 cores in total) and 64 Gb RAM.
 
@@
 
-\begin{table}
+\begin{table}[ht]
 	\caption{Comparison of local tuning strategies}
@@
 
-\begin{table}
-	\caption{Average running time $T_{av}$ of the parallel algorithm with local tuning}
-	\label{tab:3}
-	\center
-	\begin{tabular}{|c|c|c|c|}
-		\hline	
-	\diaghead{\theadfont Diag Column} { $N_{p}$}{$N_{t}$} & 1 & 4 & 8\\
-
-		\hline		
-		2  & 199.0 & 65.4 & 61.0  \\
-		5  & 91.2 & 28.0 &  26.1 \\
-		17  & 37.2 & 10.8 & 7.5 \\
-		65  & 29.3 & 8.7 & 5.4 \\
-		\hline
-	\end{tabular}
-\end{table}
+%\begin{table}[ht]
+	%\caption{Average running time $T_{av}$ of the parallel algorithm with local tuning}
+	%\label{tab:3}
+	%\center
+	%\begin{tabular}{|c|c|c|c|}
+		%\hline	
+	%\diaghead{\theadfont Diag Column} { $N_{p}$}{$N_{t}$} & 1 & 4 & 8\\	
+		%\hline		
+		%2  & 199.0 & 65.4 & 61.0  \\
+		%5  & 91.2 & 28.0 &  26.1 \\
+		%17  & 37.2 & 10.8 & 7.5 \\
+		%65  & 29.3 & 8.7 & 5.4 \\
+		%\hline
+	%\end{tabular}
+%\end{table}
 
@@
 %Russian
-Параллельный алгоритм был применен для решения той же серии из 100 problems with 5 discrete and 4 continuous parameters.
+Параллельный алгоритм был применен с ранее использованными параметрами для решения той же серии из 100 problems with 5 discrete and 4 continuous parameters.
+Fig.~ \ref{fig:3} shows the speedup $S(p)$ depending on the involved number of processes and threads in each process.
+Под ускореним здесь понимается отношение времени работы последовательного алгоритма $t_1$ ко времени работы параллельного $t_p$, т.е.
+$S(p) = t_1/t_p$, где $p$ -- суммарное число использованных threads.
 
-Table \ref{tab:3} shows the average solution time  $T_{av}$ (in seconds) depending on the involved number of processes $N_{p}$ and threads $N_{t}$ in each process.
-Our experiments involved from 1 to 16 cluster nodes depending on the number of parallel processes.
-The acceleration provided by the parallel algorithm is not ideal, but quite acceptable.
+Our experiments involved from 1 to 16 cluster nodes depending on the number of parallel processes. The speedup provided by the parallel algorithm is not ideal, but quite acceptable.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.6\textwidth]{speedup.pdf}
+    \caption{Speedup of the parallel algorithm with local tuning}
+    \label{fig:3}
+\end{figure}
 
@@
 These problems can also include non-convex constraints, which are accounted for using a special index scheme.
-
-%Russian
-Основным предположением, которое обеспечивает применимость предложенного алгоритма, является выполнение условия Липшица для функций задачи при варьировании непрерывных параметров для каждого фиксированного набора дискретных параметров.
-Задачи оптимизации, в которых функции являются липшицевыми, относятся к классу вычислительно-трудоемких задач.
-Любой численный алгоритм решения такой задачи требует вычисления значений функции в узлах некоторой (равномерной или неравномерной) сетки во всей области допустимых изменений параметров.
-Несмотря на то, что предположение липшицевости позволяет провести оценку оптимума на существенно неравномерной (т.е.
-более экономной по сравнению с равномерной) сетке в области поиска, масштаб решаемых задач здесь ограничен.
-За приемлемое время даже с использованием параллельного алгоритма можно решать задачи, в которых порядка 10 непрерывных и 10 дискретных параметров.
-Решение задач с сотнями переменных может быть основано на схеме вложенной оптимизации в предположении, что зависимость целевой функции от части переменных носит локальный характер \cite{Barkalov2020_1}.
 
@@
 %Russian
+Основным предположением, которое обеспечивает применимость предложенного алгоритма, является выполнение условия Липшица для функций задачи при варьировании непрерывных параметров для каждого фиксированного набора дискретных параметров.
+Задачи оптимизации, в которых функции являются липшицевыми, относятся к классу вычислительно-трудоемких задач.
+Любой численный алгоритм решения такой задачи требует вычисления значений функции в узлах некоторой (равномерной или неравномерной) сетки во всей области допустимых изменений параметров.
+Несмотря на то, что предположение липшицевости позволяет провести оценку оптимума на существенно неравномерной (т.е.
+более экономной по сравнению с равномерной) сетке в области поиска, масштаб решаемых задач здесь ограничен по сравнению с задачами на локальный экстремум.
+За приемлемое время даже с использованием параллельных алгоритмов можно решать задачи, в которых порядка 10 непрерывных и 10 дискретных параметров.
+Решение задач с сотнями переменных может быть основано на схеме вложенной оптимизации в предположении, что зависимость целевой функции от части переменных носит локальный характер \cite{Barkalov2020_1}.
+
 Примером сложной прикладной задачи, к которой планируется применить разработанные алгоритмы, является выбор оптимальных параметров конструкции aircraft hot-air anti-icing system для самолетов российского производства.

### 2021-03-21 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
after corrections

--- 2021-02-18 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
@@
 The PMGAL algorithm proposed in this work (in its sequential version) was compared with the genetic algorithm for solving mixed-integer optimization problems implemented in the Matlab Global Optimization Toolbox.
-Table \ref{tab:1} shows the number of trials it took for these methods to solve a number of mixed-integer test problems taken from \cite{Deep,Floudas}.
+Table \ref{tab:1} shows the number of trials and time it took for these methods to solve a number of mixed-integer test problems taken from \cite{Deep,Floudas}.
+%Russian
+The test problems used contained 2 continuous and 1 discrete parameter, which had from 3 to 6 values, depending on the specific problem.
+Also, the problem statement included from 1 to 4 non-convex constraints.  
+Note that the presence of complex constraints that form a non-convex feasible region prevents the use of methods based on the principles of local optimization. 
+The genetic algorithm from Matlab Global Optimization Toolbox was used with the recommended parameters (for example, the population size was 50, the maximum number of generations was 200, and the crossover fraction was 0.8).
+The PMGAL algorithm was used with standard parameters providing convergence to the solution ($r_\nu$ from (\ref{R}) was equal to 2.3 for all $\nu = 1,2,3$, the evolvent construction parameter $M=10$).
 The same accuracy of search $\epsilon = 10^{-2}$ was used for both methods.
+
 All computational experiments were carried out on a computer with an Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM.
@@
 
-\begin{table}
-	\caption{Average running time $T_{av}$ of the parallel algorithm with local tuning}
-	\label{tab:3}
-	\center
-	\begin{tabular}{|c|c|c|c|}
-		\hline	
-	\diaghead{\theadfont Diag Column} { $N_{p}$}{$N_{t}$} & 1 & 4 & 8\\
+%\begin{table}[ht]
+	%\caption{Average running time $T_{av}$ of the parallel algorithm with local tuning}
+	%\label{tab:3}
+	%\center
+	%\begin{tabular}{|c|c|c|c|}
+		%\hline	
+	%\diaghead{\theadfont Diag Column} { $N_{p}$}{$N_{t}$} & 1 & 4 & 8\\	
+		%\hline		
+		%2  & 199.0 & 65.4 & 61.0  \\
+		%5  & 91.2 & 28.0 &  26.1 \\
+		%17  & 37.2 & 10.8 & 7.5 \\
+		%65  & 29.3 & 8.7 & 5.4 \\
+		%\hline
+	%\end{tabular}
+%\end{table}
 
-		\hline		
-		2  & 199.0 & 65.4 & 61.0  \\
-		5  & 91.2 & 28.0 &  26.1 \\
-		17  & 37.2 & 10.8 & 7.5 \\
-		65  & 29.3 & 8.7 & 5.4 \\
-		\hline
-	\end{tabular}
-\end{table}
+The last experiment was performed to evaluate the acceleration of the parallel version of the PMGAL algorithm. 
+%Russian
+We used the parallel algorithm with previously used parameters to solve the same series of 100 problems with 5 discrete and 4 continuous parameters.
+Fig.~\ref{fig:3} shows the speedup $S(p)$ depending on the number of processes and threads involved in each process.
+Here, the acceleration is the ratio of the running time of the sequential algorithm  $t_1$ to the running time of the parallel algorithm $t_p$, i.e.
+$S(p) = t_1/t_p$, where $p$ is the total number of threads used.
+Our experiments involved from 1 to 16 cluster nodes depending on the number of parallel processes. The speedup provided by the parallel algorithm is not ideal, but quite acceptable.
 
-The last series of experiments was performed to evaluate the acceleration of the parallel version of the PMGAL algorithm. 
-Table \ref{tab:3} shows the average solution time  $T_{av}$ (in seconds) depending on the involved number of processes $N_{p}$ and threads $N_{t}$ in each process.
-Our experiments involved from 1 to 16 cluster nodes depending on the number of parallel processes.
-The acceleration provided by the parallel algorithm is not ideal, but quite acceptable.
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.6\textwidth]{speedup.pdf}
+    \caption{Speedup of the parallel algorithm with local tuning}
+    \label{fig:3}
+\end{figure}
 
@@
 These problems can also include non-convex constraints, which are accounted for using a special index scheme.
+
 We propose a parallel algorithm for solving the problems of this class.
 The basic principle of organizing parallel computations in the algorithm is based on the simultaneous carrying out of several trials at different points of the search domain.
-
 The parallel global optimization algorithm uses a local tuning technique based on the assumption that the multiextremity of the problem is weak.
 This assumption is true, in particular, at the final phase of the search, when one or more trials have already been performed in the vicinity of the global solution, and the only remaining step is to refine the found estimate of the solution.
-
 In its sequential form, the proposed algorithm remains on par with the similar methods implemented in Matlab Global Optimization Toolbox. 
 In a parallel form, the proposed algorithm demonstrates acceptable scalability up to dozens of parallel processes.
+
+%Russian
+The principal assumption which ensures that the proposed algorithm is applicable is that the Lipschitz condition for the functions of the problem with varying continuous parameters for each fixed set of discrete parameters is fulfilled.
+Optimization problems in which the functions are Lipschitzian belong to the class of computationally expensive problems.
+Any numerical algorithm for solving such a problem requires the values of the function to be calculated at the nodes of a particular (uniform or non-uniform) grid in the entire domain of admissible parameter changes.
+Although the assumption of Lipschitzness makes it possible to estimate the optimum on a substantially non-uniform (i.e., more resource-efficient than uniform) grid in the search domain, the scale of the problems being solved here is rather limited in comparison with the problems for a local extremum.
+In a reasonable time, even using parallel algorithms, it is possible to solve problems with approximately 10 continuous and 10 discrete parameters.
+A solution to problems with hundreds of variables can be based on a nested optimization scheme assuming that the dependence of the objective function on some of the variables is local \cite{Barkalov2020_1}.
+
+An example of a complex applied problem to which the developed algorithms could be applied is the selection of the optimal design parameters of an aircraft hot-air anti-icing system.
+The mathematical model of the anti-icing system, along with the parameters that change continuously (for example, the diameter of the tube supplying warm air from the engine to the wing; the diameter of the outlets on this tube; the distance between the outlets; etc.), also contains discrete parameters (for example, the number of rows of outlets, which due to the technical limitations may be equal to 1, 2 or 3).
+An example of such a problem is featured in \cite{Habashi}.
+The use of the algorithm proposed in the paper makes it possible to avoid optimization for each value of a discrete parameter separately, and instead cut off unpromising options in advance during the search process.
 

### 2021-03-22 2021/Lobachevskii_BarkalovLebedev/Barkalov.tex
replace some figures in black-and-white

--- 2021-03-21 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
@@
 \email[E-mail: ]{konstantin.barkalov@itmm.unn.ru}
-\affiliation{Lobachevsky State University of Nizhni Novgorod, Gagarin ave. 23, Nizhni Novgorod, 603950 Russia}
+\affiliation{Lobachevskii State University of Nizhni Novgorod,
+Gagarin ave. 23, Nizhni Novgorod, 603950 Russia}
 
 \author{\firstname{V.~P.}~\surname{Gergel}}
-\email[E-mail: ]{gergel@unn.ru}
-\affiliation{Lobachevsky State University of Nizhni Novgorod, Gagarin ave. 23, Nizhni Novgorod, 603950 Russia}
+\email[E-mail: ]{gergel@unn.ru} \affiliation{Lobachevskii State
+University of Nizhni Novgorod, Gagarin ave. 23, Nizhni Novgorod,
+603950 Russia}
 
 \author{\firstname{I.~G.}~\surname{Lebedev}}
-\email[E-mail: ]{ilya.lebedev@itmm.unn.ru}
-\affiliation{Lobachevsky State University of Nizhni Novgorod, Gagarin ave. 23, Nizhni Novgorod, 603950 Russia}
+\email[E-mail: ]{ilya.lebedev@itmm.unn.ru} \affiliation{Lobachevskii
+State University of Nizhni Novgorod, Gagarin ave. 23, Nizhni
+Novgorod, 603950 Russia}
 %\noaffiliation % If the author does not specify a place of work.
@@
 
-\received{February 20, 2021} % The date of receipt to the editor, i.e. December 06, 2017
+\received{February 20, 2021;  revised March 02, 2021; accepted March
+22, 2021}
 
 \begin{abstract} % You shouldn't use formulas and citations in the abstract.
-In this paper, we consider mixed-integer global optimization problems and propose a parallel algorithm for solving problems of this class based on information-statistical approach for solving continuous global optimization problems.
-Within this algorithm, we suggest using a local tuning scheme based on the assumption that the multiextremality of the discussed problem is weak.
-We also compare the sequential version of the algorithm with other similar methods.
-The effectiveness of parallelizing the algorithm has been confirmed by solving a series of mixed-integer global optimization problems on the Lobachevsky supercomputer.
+In this paper, we consider mixed-integer global optimization
+problems and propose a parallel algorithm for solving problems of
+this class based on information-statistical approach for solving
+continuous global optimization problems. Within this algorithm, we
+suggest using a local tuning scheme based on the assumption that the
+multiextremality of the discussed problem is weak. We also compare
+the sequential version of the algorithm with other similar methods.
+The effectiveness of parallelizing the algorithm has been confirmed
+by solving a series of mixed-integer global optimization problems on
+the Lobachevskii supercomputer.
 \end{abstract}
@@
 
-In this work, we will use an approach to reducing multidimensional problems to one-dimensional problems based on space-filling curves.
-Using mappings of the Peano-Hilbert curve $y(x)$ type, which project the segment $[0,1]$ onto the $N$-dimensional domain (\ref{D}),
-we can reduce the multidimensional conditional minimization problem in the domain $D$ to a one-dimensional conditional minimization problem in the segment $[0,1]$
+In this work, we will use an approach to reducing multidimensional
+problems to one-dimensional problems based on space-filling curves.
+Using mappings of the Peano--Hilbert curve $y(x)$ type, which
+project the segment $[0,1]$ onto the $N$-dimensional domain
+(\ref{D}), we can reduce the multidimensional conditional
+minimization problem in the domain $D$ to a one-dimensional
+conditional minimization problem in the segment $[0,1]$
 \begin{equation}\label{problem1}
@@
 \]
-
-These conditions allow for the classification of points $x \in [0,1]$ in accordance with the number of constraints $\nu(x)$  calculated at this point.
-The \textit{index} $\nu(x)$ can also be determined using conditions
-
+These conditions allow for the classification of points $x \in
+[0,1]$ in accordance with the number of constraints $\nu(x)$
+calculated at this point. The \textit{index} $\nu(x)$ can also be
+determined using conditions
 \begin{equation}\label{nu}
@@
 \end{gather*}
-
 \end{enumerate}
 
-The algorithm stops if the condition $\Delta_{t_j}\leq \epsilon$ is satisfied at least for one of the numbers $t_j, 1\leq j \leq p$; here $\epsilon>0$ is the target accuracy of the solution.
+The algorithm stops if the condition $\Delta_{t_j}\leq \epsilon$ is
+satisfied at least for one of the numbers $t_j, 1\leq j \leq p$;
+here $\epsilon>0$ is the target accuracy of the solution.
 
-This way of organizing parallel computing has the following rationale.
-The characteristics of intervals (\ref{R}) used in the computational rules of the algorithm can be considered as some estimates of the probability of detecting a global minimum point in these intervals.
-Inequalities (\ref{Eq:23}) order the intervals according to their characteristics, and trials are carried out in parallel in the first $p$ intervals with the highest probabilities.
+This way of organizing parallel computing has the  following
+rationale. The characteristics of intervals (\ref{R}) used in the
+computational rules of the algorithm can be considered as some
+estimates of the probability of detecting a global minimum point in
+these intervals. Inequalities (\ref{Eq:23}) order the intervals
+according to their characteristics, and trials are carried out in
+parallel in the first $p$ intervals with the highest probabilities.
 
@@
 New formulas for calculating the characteristics of search intervals $R'(i)$, which should be used at the 5th step of the algorithm, can be written as
-\[
+$
 R'(i) = h_i R(i),
-\]
+$
 where $R(i)$ from (\ref{R}), and a coefficient $h_i$ is calculated  according to the rule
@@
 
-Using the dimensionality reduction scheme by the Peano-Hilbert curve $y(x), x\in [0,1],$ we can associate each minimization problem in $y$ with a one-dimensional minimization problem
+Using the dimensionality reduction scheme by the Peano--Hilbert
+curve $y(x), x\in [0,1],$ we can associate each minimization problem
+in $y$ with a one-dimensional minimization problem
 \[
@@
 \]
-which have jump discontinuities at integer points $x_k = i, 1\leq i \leq S-1$.
-The values $z_k = g_\nu(y(x_k))$ from (\ref{trial_result}) at these points will be considered undefined, and the values of the index -- equal to 0, i.e. $\nu(x_k) = 0$ .
-
-Using the above assignments, we can reformulate the original problem as
+which have jump discontinuities at integer points $x_k = i, 1\leq i
+\leq S-1$. The values $z_k = g_\nu(y(x_k))$ from
+(\ref{trial_result}) at these points will be considered undefined,
+and the values of the index -- equal to 0, i.e. $\nu(x_k) = 0$ .
+Using the above assignments, we can reformulate the original problem
+as
 \begin{equation}\label{problem_is1}
@@
 \bibitem{Boukouvala}
-F.~Boukouvala, R.~Misener and C.~A.~Floudas, \textit{Global optimization advances in Mixed-Integer Nonlinear Programming, MINLP, and Constrained Derivative-Free Optimization, CDFO}, European J.
-Oper.
-Res.
-\textbf{252}, 701--727 (2016).
+F.~Boukouvala, R.~Misener, and C.~A.~Floudas, \textit{Global
+optimization advances in Mixed-Integer Nonlinear Programming, MINLP,
+and Constrained Derivative-Free Optimization, CDFO}, European J.
+Oper. Res. \textbf{252}, 701--727 (2016).
 
 \bibitem{Belotti}
-P.~Belotti, J.~Lee, L.~Liberti, F.~Margot and A.~W\"achter, \textit{Branching and bounds tightening techniques for non-convex MINLP}, Optim. Method. Softw. \textbf{24}(4-5), 597--634 (2009).
+P.~Belotti, J.~Lee, L.~Liberti, F.~Margot, and A.~W\"achter,
+\textit{Branching and bounds tightening techniques for non-convex
+MINLP}, Optim. Method. Softw. \textbf{24}(4-5), 597--634 (2009).
 
@@
 \bibitem{Deep}
-K.~Deep, K.~P.~Singh, M.~L.~Kansal and C.~Mohan, \textit{A real coded genetic algorithm for solving integer and mixed integer optimization problems}, Appl.
-Math.
-Comput.
-\textbf{212}(2), 505--518 (2009).
+K.~Deep, K.~P.~Singh, M.~L.~Kansal, and C.~Mohan, \textit{A real
+coded genetic algorithm for solving integer and mixed integer
+optimization problems}, Appl. Math. Comput. \textbf{212}(2),
+505--518 (2009).
 
 \bibitem{Schluter}
-M.~Schl\"uter, J.~A.~Egea and J.~R.~Banga, \textit{Extended ant colony optimization for non-convex mixed integer nonlinear programming}, Computers and Operations Research \textbf{36}(7), 2217--2229 (2009).
+M.~Schl\"uter, J.~A.~Egea, and J.~R.~Banga, \textit{Extended ant
+colony optimization for non-convex mixed integer nonlinear
+programming}, Computers and Operations Research \textbf{36}(7),
+2217--2229 (2009).
 
@@
 \bibitem{Strongin2013}
-R.~G.~Strongin, V.~P.~Gergel, V.~A.~Grishagin and K.~A.~Barkalov, \textit{Parallel computations for global optimization problems} (Moscow State University, Moscow, 2013) [In Russian].
+R.~G.~Strongin, V.~P.~Gergel, V.~A.~Grishagin, and K.~A.~Barkalov,
+\textit{Parallel computations for global optimization problems}
+(Moscow State University, Moscow, 2013) [In Russian].
 
 \bibitem{Evtushenko2009}
-Yu.~G.~Evtushenko, V.~U.~Malkova and A.~A.~Stanevichyus, \textit{Parallel global optimization of functions of several variables}, Computational Mathematics and Mathematical Physics \textbf{49}(2), 246--260 (2009).
+Yu.~G.~Evtushenko, V.~U.~Malkova, and A.~A.~Stanevichyus,
+\textit{Parallel global optimization of functions of several
+variables}, Computational Mathematics and Mathematical Physics
+\textbf{49}(2), 246--260 (2009).
 
 \bibitem{Zilinskas2011}
-R.~Paulavi\v{c}ius, J.~\v{Z}ilinskas and A.~Grothey, \textit{Parallel branch and bound for global optimization with combination of Lipschitz bounds}, Optimization Methods \& Software \textbf{26}(3), 487--498 (2011).
+R.~Paulavi\v{c}ius, J.~\v{Z}ilinskas, and A.~Grothey,
+\textit{Parallel branch and bound for global optimization with
+combination of Lipschitz bounds}, Optimization Methods \& Software
+\textbf{26}(3), 487--498 (2011).
 
@@
 \bibitem{Sergeyev2013}
-Ya.~D.~Sergeyev, R.~G.~Strongin and D.~Lera, \textit{Introduction to Global Optimization Exploiting Space-Filling Curves} (Springer Briefs in Optimization, Springer, 2013).
+Ya.~D.~Sergeyev, R.~G.~Strongin, and D.~Lera, \textit{Introduction
+to Global Optimization Exploiting Space-Filling Curves} (Springer
+Briefs in Optimization, Springer, 2013).
 
 \bibitem{Strongin2020}
-R.~Strongin, K.~Barkalov and S.~Bevzuk, \textit{Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints}, Soft Computing \textbf{24}(16), 11853--11865 (2020).
+R.~Strongin, K.~Barkalov, and S.~Bevzuk, \textit{Global optimization
+method with dual Lipschitz constant estimates for problems with
+non-convex constraints}, Soft Computing \textbf{24}(16),
+11853--11865 (2020).
 
 \bibitem{Gergel2020}
-V.~Gergel, E.~Kozinov and K.~ Barkalov, \textit{Computationally efficient approach for solving lexicographic multicriteria optimization problems}, Optim. Lett. (2020). DOI: 10.1007/s11590-020-01668-y
+V.~Gergel, E.~Kozinov, and K.~ Barkalov, \textit{Computationally
+efficient approach for solving lexicographic multicriteria
+optimization problems}, Optim. Lett. (2020). DOI:
+10.1007/s11590-020-01668-y
 
 \bibitem{Sergeyev2003}
-Y.~D.~Sergeyev, P.~Pugliese and D.~Famularo, \textit{Index information algorithm with local tuning for solving multidimensional global optimization problems with multiextremal constraints}, Math.
-Program.
-\textbf{96}(3), 489--512 (2003).
-2003.
+Y.~D.~Sergeyev, P.~Pugliese, and D.~Famularo, \textit{Index
+information algorithm with local tuning for solving multidimensional
+global optimization problems with multiextremal constraints}, Math.
+Program. \textbf{96}(3), 489--512 (2003). 2003.
 
 \bibitem{Sergeyev2007}
-Y.~D.~Sergeyev, D.~E.~Kvasov and F.~M.~H.~Khalaf, \textit{A one-dimensional local tuning algorithm for solving GO problems with partially defined constraints} Optimization Letters \textbf{1}(1), 85--99 (2007).
+Y.~D.~Sergeyev, D.~E.~Kvasov, and F.~M.~H.~Khalaf, \textit{A
+one-dimensional local tuning algorithm for solving GO problems with
+partially defined constraints} Optimization Letters \textbf{1}(1),
+85--99 (2007).
 
 \bibitem{Sergeyev2020}
-D.~E.~Kvasov, M.~S.~Mukhametzhanov, M.~C.~Nasso and Y.~D.~Sergeyev, \textit{On acceleration of derivative-free univariate Lipschitz global optimization methods}, Lecture Notes in Computer Science \textbf{11974}, 413--421 (2020).
+D.~E.~Kvasov, M.~S.~Mukhametzhanov, M.~C.~Nasso, and Y.~D.~Sergeyev,
+\textit{On acceleration of derivative-free univariate Lipschitz
+global optimization methods}, Lecture Notes in Computer Science
+\textbf{11974}, 413--421 (2020).
 
 \bibitem{Barkalov2010}
-K.~Barkalov, V.~Ryabov, S.~Sidorov, \textit{Parallel scalable algorithms with mixed local-global strategy for global optimization problems}, Lecture Notes in Computer Science \textbf{6083}, 232–240 (2010).
+K.~Barkalov, V.~Ryabov, and S.~Sidorov, \textit{Parallel scalable
+algorithms with mixed local-global strategy for global optimization
+problems}, Lecture Notes in Computer Science \textbf{6083}, 232--240
+(2010).
 
 \bibitem{Strongin2018}
-R.~G.~Strongin, V.~P.~Gergel, K.~A.~Barkalov and A.~V.~Sysoyev, \textit{Generalized parallel computational schemes for time-consuming global optimization}, Lobachevskii Journal of Mathematics \textbf{39}(4), 576--586 (2018.)
+R.~G.~Strongin, V.~P.~Gergel, K.~A.~Barkalov, and A.~V.~Sysoyev,
+\textit{Generalized parallel computational schemes for
+time-consuming global optimization}, Lobachevskii Journal of
+Mathematics \textbf{39}(4), 576--586 (2018.)
 
 \bibitem{Barkalov2020}
-R.~G.~Strongin, V.~P.~Gergel and K.~A.~Barkalov, \textit{Adaptive global optimization based on a block-recursive dimensionality reduction scheme}, Automation and Remote Control \textbf{81}(8), 1475--1485 (2020).
+R.~G.~Strongin, V.~P.~Gergel, and K.~A.~Barkalov, \textit{Adaptive
+global optimization based on a block-recursive dimensionality
+reduction scheme}, Automation and Remote Control \textbf{81}(8),
+1475--1485 (2020).
 
@@
 \bibitem{Gaviano}
-M.~Gaviano, D.~Lera, D.~E.~Kvasov and Y.~D.~Sergeyev, \textit{Software for generation of classes of test functions with known local and global minima for global optimization}, ACM Trans.
-Math.
-Softw.
-\textbf{29}, 469--480 (2003).
+M.~Gaviano, D.~Lera, D.~E.~Kvasov, and Y.~D.~Sergeyev,
+\textit{Software for generation of classes of test functions with
+known local and global minima for global optimization}, ACM Trans.
+Math. Softw. \textbf{29}, 469--480 (2003).
 
 \bibitem{Barkalov2020_1}
-K.~Barkalov, I.~Lebedev, M.~Kocheganova, V.~Gergel, \textit{Combining local and global search in a parallel nested optimization scheme}, Communications in Computer and Information Science, \textbf{1263}, 100-112 (2020).
+K.~Barkalov, I.~Lebedev, M.~Kocheganova, and V.~Gergel,
+\textit{Combining local and global search in a parallel nested
+optimization scheme}, Communications in Computer and Information
+Science \textbf{1263}, 100--112 (2020).
 
 \bibitem{Habashi}
-M.~P.~C. Pellissier, W.~G. Habashi, A.~Pueyo. \textit{Optimization via FENSAP-ICE of aircraft hot-air anti-icing systems}, Journal of Aircraft \textbf{48}(1), 265--276 (2011).
-
-%\bibitem{Floudas}
-%C.~A.~Floudas and M.~P.~Pardalos, \textit{Recent advances in global optimization} (Princeton University Press, 2016).
-%
-%\bibitem{Locatelli}
-%M.~Locatelli and F.~Schoen, \textit{Global optimization: theory, algorithms and applications} (SIAM, 2013).
-%
-%
-%\bibitem{Pardalos}
-%P.~M.~Pardalos, A.~A.~Zhigljavsky and J.~\v{Z}ilinskas \textit{Advances in stochastic and deterministic global optimization} (Springer, 2016).
-%
-%
-%\bibitem{Ciegis}
-%R.~\v{C}iegis, D.~Henty, B.~K\r{a}gstr\"om and J.~\v{Z}ilinskas, \textit{Parallel scientific computing and optimization: advances and applications}  (Springer, 2009). 
-%
-%\bibitem{Luque}
-%G.~Luque and E.~Alba, \textit{Parallel genetic algorithms. Theory and real world applications} (Springer-Verlag, Berlin, 2011).
-%
-%\bibitem{Strongin3}
-%R.~G.~Strongin, \textit{Numerical Methods in Multiextremal Problems (Information-Statistical Algorithms)} (Nauka, Moscow, 1978) [In Russian].
-%
-%\bibitem{Strongin4}
-%R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global multidimensional optimization on parallel computer}, Parallel Computing \textbf{18} (11), 1259--1273 (1992).
-%
-%\bibitem{Sergeyev4}
-%Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel asynchronous global search and the nested optimization scheme}, J. Comput. Anal. Appl., \textbf{3} (2), 123--145 (2001).
-%
-%\bibitem{Gergel1}
-%V.~P.~Gergel and S.~V.~Sidorov, \textit{A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems}, Lecture Notes in Computer Science \textbf{9251}, 505--515 (2015).
-%
-%\bibitem{Gergel2}
-%V.~Gergel, \textit{An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems}, in Proceedings of the Second International Conference on Mathematics and Computers in Sciences and in Industry (MCSI), Sliema, 13--18 (2015).
-%
-%\bibitem{Barkalov}
-%K.~Barkalov, V.~Gergel and I.~Lebedev, \textit{Solving global optimization problems on GPU cluster}, AIP Conference Proceedings \textbf{1738}, 400006 (2016).
-%
-%\bibitem{Gergel3}
-%V.~Gergel and E.~Kozinov, \textit{Efficient methods of multicriterial optimization based on the intensive use of search information}, Springer Proceedings in Mathematics and Statistics \textbf{197}, 27--45 (2017).
-%
-%\bibitem{Gergel4}
-%V.~Gergel and E.~Kozinov, \textit{Parallel computing for time-consuming multicriterial optimization problems}, Lecture Notes in Computer Science \textbf{10421}, 446--458 (2017).
-%
-%\bibitem{Gergel5}
-%V.~Gergel, V.~Grishagin and A.~Gergel, \textit{Adaptive nested optimization scheme for multidimensional global search}, J. Glob. Optim.  \textbf{66} (1), 35--51 (2016).
-%
-%\bibitem{Lera}
-%D.~Lera, Y.~D.~Sergeyev, \textit{Lipschitz and Holder global optimization using space-filling curves},  Appl. Numer. Math. \textbf{60} (1-2), 115--129 (2010).
-%
-%\bibitem{Grishagin1}
-%V~A.~Grishagin.
-\textit{On convergence conditions for a class of global search algorithms}, in Proceedings of the 3-rd All-Union seminar ``Numerical methods of nonlinear programming'', Kharkov, 82--84 (1979) [In Russian].
-%
-%\bibitem{Grishagin2}
-%V.~A.~Grishagin, Y.~D.~Sergeyev and R.~G.~Strongin, \textit{Parallel characteristic algorithms for solving problems of global optimization}, J. Glob. Optim. \textbf{10} (2), 185--206 (1997).
-%
-%\bibitem{Strongin5}
-%R.~G.~Strongin, \textit{Algorithms for Multi-extremal Mathematical Programming Problems Employing the Set of Joint Space-filling Curves}, J. Glob. Optim. \textbf{2} (4), 357--378 (1992).
-%
-%\bibitem{Sysoyev}
-%A.~Sysoyev, K.~Barkalov, V.~Sovrasov, I.~Lebedev and V.~Gergel, \textit{Globalizer -- A parallel software system for solving global optimization problems}, Lecture Notes in Computer Science \textbf{10421}, 492--499 (2017).
-%
-%\bibitem{Sergeyev7}
-%Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel Asynchronous Global Search and the Nested Optimization Scheme}, J. Comput. Anal. Appl. \textbf{3} (2), 123--145 (2001).
-%
-%\bibitem{Modorskii}
-%V.~Y.~Modorskii, D.~F.~Gaynutdinova, V.~P.~Gergel and K.~A.~Barkalov, \textit{Optimization in design of scientific products for purposes of cavitation problems}, AIP Conference Proceedings \textbf{1738}, 400013 (2016).
-%
-%\bibitem{Gergel6}
-%V.~P.~Gergel, M.~I.~Kuzmin, N.~A.~Solovyov and V.~A.~Grishagin, \textit{Recognition of surface defects of cold-rolling sheets based on method of localities}, International Review of Automatic Control \textbf{8} (1), 51--55 (2015).
+M.~P.~C. Pellissier, W.~G. Habashi, and A.~Pueyo.
+\textit{Optimization via FENSAP-ICE of aircraft hot-air anti-icing
+systems}, J. of Aircraft \textbf{48}(1), 265--276 (2011).
 

### 2021-03-26 2021/PСT_GergelKozinov/2021_PCT.tex
PCT финальная версия

@@
+\makeatletter\@ifundefined{UseRawInputEncoding}{\newcommand\UseRawInputEncoding{}}{\UseRawInputEncoding}\makeatother
 \documentclass{svproc}
@@
 \begin{abstract}
-The paper discusses a new approach to solving computationally time-consuming multicriteria optimization problems, in which some variable parameters can only take on discrete values.
+The paper discusses a new approach to solving computationally time-consuming multicriteria optimization problems in which some variable parameters can only take on discrete values.
 Under the proposed approach, the solution of mixed-integer optimization problems is reduced to solving a family of optimization problems where only continuous parameters are used.
-All problems of the family are solved simultaneously in the time-shared mode, where the optimization problem for the next global search iteration is selected adaptively taking into account the search information obtained in the course of the calculations.
-The algorithms proposed enable parallel computing on high-performance computing systems.
-Computational experiments confirm that the proposed approach makes it possible to significantly reduce computation volume and time for solving complex multicriteria mixed-integer optimization problems.
-\keywords{Multicriteria optimization $\cdot$ mixed-integer optimization problems $\cdot$ methods of criteria scalarization $\cdot$ global optimization $\cdot$ search information $\cdot$ parallel computing $\cdot$ computational experiment}
+All problems of the family are solved simultaneously in time-shared mode, where the optimization problem for the next global search iteration is selected adaptively, taking into account the search information obtained in the course of the calculations.
+The suggested algorithms enable parallel computing on high-performance computing systems.
+The computational experiments confirm that the proposed approach can significantly reduce the computation volume and time required for solving complex multicriteria mixed-integer optimization problems.
+
+\keywords{Multicriteria optimization $\cdot$ Mixed-integer optimization problems $\cdot$ Methods of criteria scalarization $\cdot$ Global optimization $\cdot$ Search information $\cdot$ Parallel computing $\cdot$ Computational experiment}
 \end{abstract}
@@
 
-Multicriteria optimization (MCO) problems arise almost every time when the optimal decision need to be chosen when developing complex technical devices and systems.
-The scale of demand for MCO problems determines the high intensity of research in this area -- see, for example, monographs \cite{c1,c2,c3,c4,c5} and reviews of scientific and practical results in this area \cite{c6,c7,c8,c9}.
+Multicriteria optimization (MCO) problems arise whenever optimal decisions need to be made during the development of complex technical devices and systems.
+The scale of demand for MCO problems determines the high intensity of research in this area (see, for example, the monographs \cite{c1,c2,c3,c4,c5} and reviews of scientific and practical results in this area \cite{c6,c7,c8,c9}).
 
-Usually, the solution of a MCO problem is a set of efficient (non-dominated) decisions for which an improvement in values with respect to some criteria cannot be achieved without deteriorating the efficiency with respect to some other criteria.
-The definition of the whole set of efficient decisions (the Pareto set), on the one hand, may require a large amount of computation and, on the other hand, may be redundant as the analysis of a large number of efficient decisions may require a significant effort for the person who makes the decision (decision maker, DM).
+Usually, the solution to an MCO problem is a set of efficient (non-dominated) decisions which cannot be improved under some criteria without deteriorating the efficiency under some other criteria.
+The definition of the whole set of efficient decisions (the Pareto set) may, on the one hand, require a large amount of computation and, on the other, may be redundant as the analysis of a large number of efficient decisions may require a significant effort from the person who makes the decision (the decision maker, DM).
 Thus, it may be practically justified to find only a relatively small set of efficient decisions that can be formed according to the optimality requirements defined by the DM.
 
-By limiting the set of efficient decisions to be computed, one can achieve a noticeable reduction in the amount of calculations required.
+By restricting the set of efficient decisions to be computed, one can achieve a noticeable reduction in the number of computations required.
 However, efficiency criteria may be complex, \textit{multiextremal}, and calculating the values of these criteria can prove \textit{computationally demanding}.
-In addition, some variables may only take on discrete values.
-In such cases, MCO problems involve \textbf{significant computational complexity}, which can only be overcome by using high-performance supercomputer systems.
+Besides, some variables may only take on discrete values.
+In such cases, MCO problems involve a \textbf{significant computational complexity}, which can only be overcome by using high-performance supercomputer systems.
 
 Many different approaches have been proposed for solving MCO problems (see, for example, \cite{c3,c6,c10,c11}.
-Most commonly, various methods based on reduction of the vector criterion to a particular scalar function are used \cite{c2,c12}.
-The number of works on multicriteria mixed-integer problems is however more limited -- in most cases, the issues of discrete parameter analysis are considered in relation to scalar optimization problems (see, for example, reviews \cite{c13,c14}).
-The widely used deterministic methods for solving problems of this class are usually based on the Branch-and-Bound \cite{c15} or on the Branch-and-Reduce approaches \cite{c16}.
-A number of meta-heuristic and genetic algorithms are also known and are based, in one way or another, on the random search concept \cite{c17,c18}.
+Most commonly, various methods based on the reduction of the vector criterion to a particular scalar function are used \cite{c2,c12}.
+The number of works concerned with multicriteria mixed-integer problems is however more limited; in most cases, the issues of discrete parameter analysis are considered in connection with scalar optimization problems (see, for example, reviews \cite{c13,c14}).
+The widely used deterministic methods for solving problems of this class are usually based on either the Branch-and-Bound \cite{c15} or the Branch-and-Reduce approaches \cite{c16}.
+Several meta-heuristic and genetic algorithms are also known and are based, in one way or another, on the random search concept \cite{c17,c18}.
 
-This paper presents the results of research that was carried out to develop highly effective parallel methods of multicriteria optimization making use of all the search information obtained in the course of calculations \cite{c19,c20,c21}.
-A new contribution to this research area consists in the development of an approach for solving MCO problems, in which some of the varied parameters can only take discrete values.
-The proposed approach reduces the solution of mixed-integer optimization problems to a family of optimization problems where only continuous parameters are used.
-All problems of the family are solved simultaneously in the time-shared mode, where the selection of the optimization problem for the next iteration of the global search is performed adaptively taking into account the search information obtained in the course of the calculations.
-The algorithms developed enable efficient parallel computing on high-performance computing systems.
+This paper presents the results of a study concerned with the development of highly effective parallel methods of multicriteria optimization making use of all the search information obtained throughout computations \cite{c19,c20,c21}.
+A new contribution to this research field is the development of an approach for solving MCO problems in which some of the varied parameters can only take on discrete values.
+The suggested approach reduces the solution of mixed-integer optimization problems to a family of optimization problems where only continuous parameters are used.
+All problems of the family are solved simultaneously in time-shared mode, where the selection of the optimization problem for the next iteration of the global search is performed adaptively, taking into account the search information obtained in the course of computations.
+The developed algorithms enable efficient parallel computing on high-performance computing systems.
 
-The rest of the paper is organized as follows.
-In Section \ref{sec:2}, we give the statement of multicriteria optimization problems, present a minimax scheme for scalarization of the vector efficiency criterion and introduce the concept of multistage solution of multicriteria optimization problems.
-In Section \ref{sec:3}, we present a description of the proposed approach based on the reduction of mixed-integer optimization problems to solving a family of optimization problems with the use of only continuous parameters.
-This section also describes a dimensionality reduction scheme, which allows multidimensional optimization problems to be reduced to a one-dimensional global search problem.
-Section \ref{sec:4} presents a parallel algorithm for solving multicriteria mixed-integer optimization problems, which is used in the framework of the proposed approach.
-Section \ref{sec:5} contains the results of numerical experiments confirming that the proposed approach is promising.
-In conclusion, the results obtained are discussed and possible main directions for continuing the research are outlined.
+The rest of the paper is structured as follows.
+In Section \ref{sec:2}, we statement of multicriteria optimization problems, present a minimax scheme for scalarization of the vector efficiency criterion, and introduce the concept of multistage solution for multicriteria optimization problems.
+In Section \ref{sec:3}, we describe the proposed approach based on the reduction of mixed-integer optimization problems to the solution of a family of optimization problems using only continuous parameters.
+In that section, we also describe a dimensionality reduction scheme by which multidimensional optimization problems can be reduced to one-dimensional global search problems.
+Section \ref{sec:4} presents a parallel algorithm for solving multicriteria mixed-integer optimization problems under of the proposed approach.
+Section \ref{sec:5} contains the results of numerical experiments that confirm that the approach we propose is promising.
+In the Conclusions section, we discuss the results obtained and outline possible directions for future research.
 
@@
 
-The problem of multicriteria mixed-integer optimization (MCOmix) can be formulated as follows
+The problem of multicriteria mixed-integer optimization (MCOmix) can be expressed as follows
 \begin{equation}\label{eq:1}
-f(y,u)\to \min, y\in D,u \in U,
+f(y,u)\to \min,\ y\in D,\ u \in U,
 \end{equation}
-where $f(y,u)=(f_1 (y,u),f_2 (y,u), \dots,f_s (y,u))$ is the vector criterion of efficiency in which the varied parameters belong to two different types:
+where $f(y,u)=(f_1 (y,u),f_2 (y,u), \dots,f_s (y,u))$ is the vector criterion of efficiency, in which the varied parameters belong to two different types:
 \begin{itemize}
-	\item continuous parameters $y=(y_1,y_2,\dots,y_n)$, whose domain of possible values is represented by an $N$-dimensional hyperparallelepiped:
+	\item continuous parameters $y=(y_1,y_2,\dots,y_n)$, whose domain of possible values is represented by an $N$-dimensional hyperparallelepiped,
 \begin{equation}\label{eq:2}
@@
 \end{equation}
-for specified vectors $a$ and $b$.
-	\item discrete parameters  $u=(u_1,u_2, \dots ,u_m)$, each of which can only take on a fixed (discrete) set of values:
+for specified vectors $a$ and $b$;
+
+	\item discrete parameters  $u=(u_1,u_2, \dots ,u_m)$, each of which can only take on a fixed (discrete) set of values,
 \begin{equation}\label{eq:3}
-U=U_1 \times U_2 \times \dots \times U_m=\{w_k=<w_{1k},w_{2k}, \dots, w_{mk}>: 1 \leq k \leq l\}, w_{ik} \in U_i, 
+U=U_1 \times U_2 \times \dots \times U_m=\{w_k=\langle w_{1k},w_{2k}, \dots, w_{mk}\rangle : 1 \leq k \leq l\},\ w_{ik} \in U_i,
 \end{equation}
-where $U_i=\{v_{i1},v_{i2}, \dots ,v_{il_i} \}$, $1 \leq i \leq m$, is a set of $l_i>0$ admissible discrete values for the parameter $u_i$, i.e., the whole set of all possible discrete parameter values $U$ contains
+where $U_i=\{v_{i1},v_{i2}, \dots ,v_{il_i} \}$, $1 \leq i \leq m$, is a set of $l_i>0$ admissible discrete values for the parameter $u_i$, i.e., the set $U$ of all possible discrete parameter values contains
 \begin{equation}\label{eq:4}
@@
 \end{equation}
-different elements (tuples $w_k$,$1 \leq k \leq l$).
-Without loss of generality, it will be assumed in further consideration that the criteria $f_i (y,u)$, $1 \leq i \leq s$, are non-negative and their reduction corresponds to an increase in the efficiency of the decisions selected.
+different elements (tuples $w_k$, $1 \leq k \leq l$).
+Without loss of generality, we assume below that the criteria $f_i(y,u)$, $1 \leq i \leq s$, are non-negative, and their reduction corresponds to an increase in the efficiency of the decisions selected.
 \end{itemize}
 
-In the most complex variant, the criteria $f_i(y,u)$, $1 \leq i \leq s$, may be multiextremal, and the procedures for calculating their values can be computationally time-consuming.
-It is also assumed that the criteria $f_i (y,u)$, $1 \leq i \leq s$, meet the Lipschitz condition
+In the most difficult case, the criteria $f_i(y,u)$, $1 \leq i \leq s$, may be multiextremal, and the procedures for calculating their values can be computationally time-consuming.
+We also assume that the criteria $f_i (y,u)$, $1 \leq i \leq s$, meet the Lipschitz condition
 \begin{equation}\label{eq:5}
@@
 
-Efficiency criteria in an MCO problem are usually controversial and the parameters $(y^*,u^*) \in D \times U$ with the optimal values for all criteria at the same time may not be available.
-In such situations, it is a common approach for MCO problems to find efficient (non-dominated) decisions, for which an improvement in the values of some criteria leads to a deterioration of efficiency indicators for other criteria.
-Obtaining the whole set of efficient decisions (the Pareto set) can require a lot of computation and, as a result, another approach is often used -- to find only a relatively small set of efficient decisions defined according to the decision maker's requirements.
+Efficiency criteria in an MCO problem are usually controversial, and there may not exist parameters $(y^*,u^*) \in D \times U$ with values simultaneously optimal for all criteria.
+In such situations, it is a common practice in MCO problems to search for efficient (non-dominated) decisions, for which an improvement in the values of some criteria leads to a deterioration of the efficiency indicators for other criteria.
+Obtaining the whole set of efficient decisions (the Pareto set) can require a lot of computation and, for this reason, a different approach is often used, namely search only for a relatively small set of efficient decisions defined according to the decision maker's requirements.
 
-A commonly used approach to obtaining efficient decisions is to transform a vector criterion into some combined scalar efficiency function\footnote{It should be noted that this approach ensures that a wide range of already existing global optimization methods can be used to solve MCO problems.}
+A commonly used approach to obtaining efficient decisions is to transform the vector criterion into some combined scalar efficiency function\footnote{It should be noted that this approach ensures that a wide range of already existing global optimization methods can be used to solve MCO problems.}
 \begin{equation}\label{eq:6}
-  \min{F(\alpha,y,u)}, y \in D, u \in U,
+  \min{F(\alpha,y,u)},\ y \in D,\ u \in U,
 \end{equation}
-where $F$ is an objective function generated by scalarization of the criteria $f_i$, $1 \leq i \leq s$, $\alpha$ is the vector of the parameters of the criteria convolution applied, while $D$ and $U$ are the domains of possible parameter values from (\ref{eq:2})-(\ref{eq:3}).
-By virtue of (\ref{eq:5}), the function $F(\alpha, y, u)$ also satisfies the Lipschitz condition with some constant $L$, i.e.
+where $F$ is an objective function generated by scalarization of the criteria $f_i$, $1 \leq i \leq s$, $\alpha$ is the vector of parameters of the criteria convolution applied, while $D$ and $U$ are the domains of possible parameter values from (\ref{eq:2})--(\ref{eq:3}).
+By virtue of (\ref{eq:5}), the function $F(\alpha, y, u)$ also satisfies the Lipschitz condition with some constant $L$, that is,
 \begin{equation}\label{eq:7}
@@
   F(\lambda,y,u)=\max{(\lambda_i f_i (y,u),1\leq i \leq s)},\\
-  \lambda=(\lambda_1,\lambda_2, \dots, \lambda_s)\in \Lambda \subset R^s: \sum_{i=1}^{s}\lambda_i=1, \lambda_i \geq 0, 1 \leq i \leq s.
+  \lambda=(\lambda_1,\lambda_2, \dots, \lambda_s)\in \Lambda \subset R^s: \sum_{i=1}^{s}\lambda_i=1, \lambda_i \geq 0,\ 1 \leq i \leq s.
 \end{matrix}
@@
 
-It should be noted that due to the possible changes in the requirements for optimality in the process of calculations it may be necessary to change the parameters of the convolution $\lambda$ from (\ref{eq:8}).
+It should be noted that, due to the possible changes in the requirements for optimality in the process of calculation, it may be necessary to change the parameters of the convolution $\lambda$ from (\ref{eq:8}).
 Such variations yield a set of scalar global optimization problems (\ref{eq:6})
@@
 which is necessary for solving the MCOmix problem.
-This set of problems can be formed sequentially in the course of calculations; the problems of the set can be solved strictly sequentially or simultaneously in time-shared mode.
-In addition, the problems of the set $\mathbb{F}_T$ can be solved in parallel using high-performance computing systems.
-The possibility of forming the set $\mathbb{F}_T$ defines a \textit{new approach to multistage solution of multicriteria optimization} (MMCO) problems -- see, for example, \cite{c22}.
+This set of problems can be formed sequentially during the calculations; the problems of the set can be solved strictly sequentially or simultaneously in time-shared mode.
+Furthermore, the problems of the set $\mathbb{F}_T$ can be solved in parallel using high-performance computing systems.
+The possibility of forming the set $\mathbb{F}_T$ determines \textit{a new approach to the multistage solution of multicriteria optimization} (MMCO) problems (see, for instance, \cite{c22}).
 
@@
 
-In the presence of discrete parameters, solving multicriteria optimization  problems becomes considerably  more complicated -- in many cases, it becomes necessary to calculate criterion values for all possible values of discrete parameters.
-The proposed approach to improving the efficiency of solving MCOmix problems is based on the following two basic ideas: unrolling mixed-integer optimization problems \cite{c23} and dimensionality reduction \cite{c24,c25}.
+The solution of multicriteria optimization  problems becomes considerably  more complicated in the presence of discrete parameters: in many cases, it is necessary to calculate the criterion values for all possible values of the discrete parameters.
+The proposed approach to improving the efficiency of the solution of MCOmix problems is based on two basic ideas: unrolling mixed-integer optimization problems \cite{c23} and dimensionality reduction \cite{c24,c25}.
 
@@
 
-To solve the global optimization problem (\ref{eq:6}), a two-stage nested optimization scheme can be used
+To solve the global optimization problem (\ref{eq:6}), a two-stage nested optimization scheme can be used:
 \begin{equation}\label{eq:11}
-\begin{matrix}
+\begin{gathered}
   F(\alpha,y^*,u^*)=\min{F(\alpha,y,u)} = \min_{y \in D} \min_{u \in U}{F(\alpha,y,u)} = \\
-  \min_{y \in D}(F(\alpha,y,w_1 ),F(\alpha,y,w_2 ), \dots, F(\alpha,y,w_l )),\\
-	w_i\in U,1 \leq i \leq l.
-\end{matrix}
+  \qquad=\min_{y \in D}\bigl(F(\alpha,y,w_1 ),F(\alpha,y,w_2 ), \dots, F(\alpha,y,w_l )\bigr),\\
+	w_i\in U,\ 1 \leq i \leq l.
+\end{gathered}
 \end{equation}
 
-In the computational scheme (\ref{eq:11}), for any value of continuous parameters $y \in D$, the values of the function $F(\alpha,y,u)$ are calculated for all possible values of discrete parameters $u \in U$.
-However, the smallest value of the function $F(\alpha,y,u)$ is achieved only with one specific value of discrete parameters $u^*$, so the calculation of function $F(\alpha,y,u)$ values for other values of discrete parameters is redundant.
-Thus, for example, for the problem
+In computational scheme (\ref{eq:11}), the values of the function $F(\alpha,y,u)$ for any value of continuous parameters $y \in D$ are computed for all possible values of the discrete parameters $u \in U$.
+However, the smallest value of the function $F(\alpha,y,u)$ is achieved only with one specific value of the discrete parameters $u^*$, so the computation of the values of the function $F(\alpha,y,u)$ for other values of the discrete parameters is redundant.
+For example, consider the problem
 \begin{equation}\label{eq:12}
- \min\{u^2(sin(x)+sin(10x/3)):x\in[2.7,7.5],u\in \{1,2\}\}
+ \min\{u^2(\sin(x)+\sin(10x/3)):x\in[2.7,7.5],\ u\in \{1,2\}\},
 \end{equation}
-having one continuous and one discrete parameter, the plots of the  function $F(\alpha,y,u)$ for different values of the discrete parameter have form shown in Fig.~\ref{fig:1} and, as one can see, there is no need to calculate the value of the function $F(\alpha,y,u)$ for $u=1$.
+which has one continuous and one discrete parameter.
+The plots of the  function $F(\alpha,y,u)$ for different values of the discrete parameter have the form shown in Figure~\ref{fig:1} and, as we can see, there is no need to compute the value of the function $F(\alpha,y,u)$ for $u=1$.
 
@@
   \includegraphics[width=0.9\linewidth]{fig1}
-  \caption{The function $F(\alpha,y,u)$ of the problem (\ref{eq:12}) plotted for different values of the discrete parameter (left, $u=1$,  right, $u=2$)}
+  \caption{The function $F(\alpha,y,u)$ from problem (\ref{eq:12}) plotted for different values of the discrete parameter (on the left, $u=1$;  on the right, $u=2$)}
   \label{fig:1}
@@
 
-It is possible to increase the efficiency and, accordingly, to reduce computational complexity when solving  the problem (\ref{eq:6}) by excluding (or at least by reducing) the calculation of the function $F(\alpha,y,u)$ for discrete parameters $u \neq u^*$.
-In some rare situations, the information about the values of the discrete parameters that do not allow the minimum value of the function $F(\alpha,y,u)$ to be reached, may be known a priori.
-In most cases, such information can be obtained only in the process of calculations based on the calculated search information.
+It is possible to increase the efficiency and, accordingly, to reduce the computational complexity when solving the problem (\ref{eq:6}), by excluding (or, at least, by reducing) the computation of the function $F(\alpha,y,u)$ for discrete parameters $u \neq u^*$.
+In some rare situations,  we may know a priori information about the values of discrete parameters that do not allow the function $F(\alpha,y,u)$ to reach a minimum value.
+In most cases, that information can be obtained only in the process of computations based on the calculated search information.
 
@@
 \begin{equation}\label{eq:13}
- \Omega_k=\{ y^i,u^i,z^i=F(\alpha,y^i,u^i ): 1\leq i\leq k \}
+ \Omega_k=\{ y^i,u^i,z^i=F(\alpha,y^i,u^i ) : 1\leq i\leq k \}
 \end{equation}
 is the search information obtained in the process of calculation after $k>1$ global search iterations.
-Then the procedure for adaptive estimation of the discrete parameter value, at which the minimum value  of the function $F(\alpha,y,u)$ is expected to be obtained, can be determined using the decision rule
+Then, the procedure for adaptive estimation of the discrete parameter value at which the function $F(\alpha,y,u)$ is expected to reach the minimum value can be determined using the decision rule
 \begin{equation}\label{eq:14}
-	\theta =\Theta(\Omega_k ),1 \leq \theta \leq l, 
+	\theta =\Theta(\Omega_k ),\ 1 \leq \theta \leq l,
 \end{equation}
- that determines the most promising value of the discrete parameter $u \in U$ at each step of the global search.
+which determines the most promising value of the discrete parameter $u \in U$ at each step of the global search.
 
@@
 
-\textit{Rule 1.} Apply the decision rule $\Theta$ from (\ref{eq:14}) to available search information $\Omega_k$ from (\ref{eq:13}),  and determine the value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.
+\textit{Rule 1.} Apply the decision rule $\Theta$ from (\ref{eq:14}) to the available search information $\Omega_k$ from (\ref{eq:13})  and determine the value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.
 
-\textit{Rule 2.} Determine the point $y \in D$ of the global search iteration to be performed with a fixed value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.
+\textit{Rule 2.} Determine the point $y \in D$ of the global search iteration that needs to be performed with a fixed value of the discrete parameter $u=w_\theta$, $w_\theta \in U$.
 
-\textit{Rule 3.} Check the stopping condition for the computation.
-If the required accuracy of the global search is not achieved, it is necessary to supplement the search information $\Omega_k$ from (\ref{eq:13}) with the results of the current iteration and continue calculations starting from Rule 1.
+\textit{Rule 3.} Check the stopping condition of the computation.
+If the required accuracy of the global search is not achieved, it is necessary to supplement the search information $\Omega_k$ from (\ref{eq:13}) with the results of the current iteration and continue the computation starting from Rule~1.
 
@@
 
-It should be noted that the computational complexity of solving global optimization problems increases exponentially with increasing dimensionality.
-The problem of increasing computational complexity was even called the ``dimensionality curse''.
-In particular, when applying the computational scheme discussed in subsection \ref{subsec:31}, it is necessary to accumulate and analyze multidimensional search information $\Omega_k$ from (\ref{eq:13}).
-This computational complexity can be significantly reduced by reducing the optimization problems to be solved by using Peano \textit{curves} or \textit{evolvents} $y(x)$, which unmbiguously and continuously map a segment [0,1] on an $n$-dimensional hyper-cube $D$ -- see, for example, \cite{c24,c25}.
-As a result of this reduction, the multidimensional global optimization problem (\ref{eq:6}) is reduced to a one-dimensional problem
+Note that the computational complexity of the solution of global optimization problems increases exponentially as the dimensionality grows.
+The problem of growth of the computational complexity was even called the ``dimensionality curse''.
+In particular, when applying the computational scheme discussed in subsection \ref{subsec:31}, it is necessary to accumulate and analyze the multidimensional search information $\Omega_k$ from (\ref{eq:13}).
+This computational complexity can be significantly reduced by reducing the optimization problems to be solved, through the use of \textit{Peano curves} or \textit{evolvents} $y(x)$ that unambiguously and continuously map the segment $[0,1]$ onto an $n$-dimensional hypercube $D$ (see, for example, \cite{c24,c25}).
+As a result, the multidimensional global optimization problem in (\ref{eq:6}) is reduced to a one-dimensional problem, namely
 \begin{equation}\label{eq:15}
- \min_{y\in D, u \in U}F(\alpha,y,u) = \min_{x \in [0,1]}(F(\alpha,y(x),w_1 ),F(\alpha,y(x),w_2 ),\dots,F(\alpha,y(x),w_l ))
+\min_{y\in D, u \in U}F(\alpha,y,u) = \min_{x \in [0,1]} \bigl(F(\alpha,y(x),w_1 ),F(\alpha,y(x),w_2 ),\dots,F(\alpha,y(x),w_l )\bigr)
 \end{equation}
 
-It should be noted that the one-dimensional functions $F(\alpha,y(x),u)$ resulting from the reduction satisfy the uniform H{\" o}lder condition, i.e.
+We should also note that one-dimensional functions $F(\alpha,y(x),u)$ resulting from the reduction satisfy a uniform H{\"o}lder condition,
 \begin{equation}\label{eq:16}
-|F(\alpha, y(x_1),u_1)-F(\alpha, y(x_2),u_2)| \leq H|x_1-x_2 |^{1/n},
+|F(\alpha, y(x_1),u_1)-F(\alpha, y(x_2),u_2)| \leq H \lvert x_1-x_2\rvert^{1/n},
 \end{equation}
-where the constant $H$ is determined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:7}) and $n$ is the dimensionality of the MCOmix problem from (\ref{eq:1}).
+where the constant $H$ is determined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:7}), and $n$ is the dimensionality of the MCOmix problem from (\ref{eq:1}).
 
-As a result of dimensionality reduction, the search information $\Omega_k$ from (\ref{eq:13}) obtained in the calculation can be presented in the form 
+As a result of the dimensionality reduction, the search information $\Omega_k$ from (\ref{eq:13}) obtained by computation can be expressed as
 \begin{equation}\label{eq:17}
-A_k=\{(x_i, u_i, z_i, f_i = f(y(x_i), u_i): 1 \leq i \leq k \},
+A_k=\{(x_i, u_i, z_i, f_i = f(y(x_i), u_i) : 1 \leq i \leq k \},
 \end{equation}
-where $x_i$, $u_i$, $1 \leq i \leq k$,  are the points of the global search iterations, $z_i$, $f_i$, $1 \leq i \leq k$,  are the values of the scalar criterion $F(\alpha,y(x),u)$ and the criteria $f_i (y)$, $1 \leq i \leq s$, calculated at the points $x_i$, $u_i$, $1 \leq i \leq k$.
-Note that the data in the set $A_k$ are arranged in ascending order\footnote{Data ordering is reflected by using a subscript.} of the points $x_i$, $1 \leq i \leq k$, i.e.
+where $x_i, u_i$, $1 \leq i \leq k$,  are the points of the global search iterations, and $z_i$, $f_i$, $1 \leq i \leq k$,  are the values of the scalar criterion $F(\alpha,y(x),u)$ and criteria the $f_i (y)$, $1 \leq i \leq s$, computed at the points $x_i$, $u_i$, $1 \leq i \leq k$.
+Note that the data in the set $A_k$ are arranged in ascending order\footnote{The order is indicated by the subscript.} of the points $x_i$, $1 \leq i \leq k$, that is,
 \begin{equation}\label{eq:18}
- x_1< x_2< \dots < x_k
+ x_1< x_2< \dots < x_k,
 \end{equation}
-for a more efficient execution of global search algorithms.
+for more efficient execution of global search algorithms.
 
-The use of dimensionality reduction makes it possible to combine (concatenate) one-dimensional functions $F(\alpha, y(x),w_i)$, $1 \leq i \leq l$, from (\ref{eq:15}) into a single one-dimensional function $\Phi(\alpha,y(x))$, defined over the segment $[0,l]$ (see Fig.~\ref{fig:1}).
-
+Dimensionality reduction makes it possible to combine (concatenate) the one-dimensional functions $F(\alpha, y(x),w_i)$, $1 \leq i \leq l$, from~(\ref{eq:15}) into a single one-dimensional function $\Phi(\alpha,y(x))$ defined over the segment $[0,l]$ (see Figure~\ref{fig:1}),
 \begin{equation}\label{eq:19}
@@
 \begin{cases}
-F(\alpha,y(x),w_1 ),x \in [0,1], \\
-F(\alpha,y(x),w_2 ),x \in (1,2], \\
-\dots \\
-F(\alpha,y(x),w_l ),x \in (l-1,l],
+F(\alpha,y(x),w_1 ),\ x \in [0,1], \\
+F(\alpha,y(x),w_2 ),\ x \in (1,2], \\
+\hdotsfor{1} \\
+F(\alpha,y(x),w_l ),\ x \in (l-1,l],
 \end{cases}
 \end{equation}
-where $l$ from (\ref{eq:4}) is the number of different variants of the values of discrete parameters $u \in U$.
-The mapping of the extended segment $[0,l]$ onto the domain $D$ of the continuous parameter values from (\ref{eq:2}) can be determined as follows
+where $l$, as expressed in (\ref{eq:4}), is the number of different variants of the values of discrete parameters $u \in U$.
+The mapping of the extended segment $[0,l]$ onto the domain $D$ of continuous parameter values from (\ref{eq:2}) can be determined as
 \begin{equation}\label{eq:20}
-Y(x)=y(x-E(x)),x\in[0,l]
+Y(x)=y(x-E(x)),\ x\in[0,l]
 \end{equation}
-($E(x)$ denotes the operation of taking the integral part of the number $x$).
-It should be noted that the function $\Phi(\alpha,y(x))$ is discontinuous at the points $x_i$, $1 \leq i \leq l-1$,  further on, the values of the function $\Phi(\alpha,y(x))$ at these points are considered undefined and are not used in the calculations.
+where $E(x)$ denotes the integral part of the number $x$.
+Note that the function $\Phi(\alpha,y(x))$ is discontinuous at the points $x_i$, $1 \leq i \leq l-1$.
+Below, we consider the values of the function $\Phi(\alpha,y(x))$ at these points undefined and will not use them in computations.
 
@@
 
-In the general case, the problem of minimizing the function $F(\lambda,y,u)$ from (\ref{eq:8}) is one of  global optimization problems.
-The solution of such problems involves the construction of grids covering the search domain $D$ -- see, for example, \cite{c24,c25,c26,c27,c28,c29,c30,c31}.
+In the general case, the problem of minimizing the function $F(\lambda,y,u)$ from (\ref{eq:8}) is one of the global optimization problems.
+The solution to such problems involves the construction of grids covering the search domain $D$ (see, for example, \cite{c24,c25,c26,c27,c28,c29,c30,c31}).
 
-The proposed approach uses the algorithm of global mixed-integer search (AGMIS) to minimize the function $F(\lambda,y,u)$.
+The proposed approach relies on the algorithm of global mixed-integer search (AGMIS) to minimize the function $F(\lambda,y,u)$.
 This algorithm expands the possibilities of multiextremal optimization methods developed within the framework of the information-statistical theory of global search \cite{c19,c20,c21,c22,c23,c32,c33,c34,c35,c36,c37,c38,c39} to minimize the reduced one-dimensional function $\Phi(\alpha,y(x))$ from (\ref{eq:19}).
 
-The AGMIS general computational scheme can be presented as follows (see also \cite{c21}).
+The AGMIS general computational scheme can be described as follows (see also \cite{c21}).
 
-At the initial iteration of AGMIS, the value of the minimized function is calculated at some arbitrary point of the interval $(0,l)$(further, obtaining the function value will be called a \textit{trial}).
-Next, let $k$, $k > 1$, global search iterations be performed.
-Selection of the trial point $(k+1)$ for the next iteration is determined by the following rules.
+At the initial iteration of AGMIS, the value of the minimized function is calculated at some arbitrary point of the interval $(0,l)$ (below, the computation of the function value is called a \textit{trial}).
+Next, assume that $k$, $k > 1$, global search iterations have already been performed.
+The selection of the trial point $(k+1)$ for the next iteration is done according to the following rules.
 
-\textit{Rule 1.} For each interval $(x_{i-1}, x_i)$,$1<i \leq k$, calculate the value $R(i)$, further referred to as the interval \textit{characteristic}.
+\textit{Rule 1.} For each interval $(x_{i-1}, x_i)$, $1<i \leq k$, compute the value $R(i)$, (below, we refer to it as the \textit{interval characteristic}).
 
-\textit{Rule 2.} Determine the interval $(x_{t-1}, x_t)$, to which the maximum characteristic
+\textit{Rule 2.} Determine the interval $(x_{t-1}, x_t)$ that the maximum characteristic
 \begin{equation}\label{eq:21}
@@
 \end{equation}
-corresponds.
+corresponds to.
 
-\textit{Rule 3.} Perform a new trial at $x^{k+1}$ in the interval $(x_{t-1}, x_t)$ with the maximum characteristic
+\textit{Rule 3.} Perform a new trial at $x^{k+1}$ in the interval $(x_{t-1}, x_t)$ with maximum characteristic
 \begin{equation}\label{eq:22}
@@
 \end{equation}
-(the values of the discrete parameters $u \in U$ are given according to (\ref{eq:19})).
+(the values of the discrete parameters $u \in U$ are given as specified in~(\ref{eq:19})).
 
- The stopping condition under which the trials are stopped is determined by the following condition
+The stopping condition under which the trials are stopped is determined as
 \begin{equation}\label{eq:23}
@@
 \end{equation}
-where $t$ is from (\ref{eq:21}), $n$ is the dimensionality of the problem being solved from (\ref{eq:1}), and $\varepsilon > 0$ is the specified accuracy of the problem solution.
-If the stopping condition is not satisfied, the number of iteration $k$ is increased by one, and a new global search iteration is executed.
+where $t$ is taken from (\ref{eq:21}), $n$ is the dimensionality of the problem being solved (see (\ref{eq:1})), and $\varepsilon > 0$ is the specified accuracy of the problem solution.
+If the stopping condition is not satisfied, the number of iterations $k$ is increased by one, and a new global search iteration is executed.
 
-   To explain the computation scheme under consideration, the following may be noted.
-The calculated characteristics $R(i)$, $1<i \leq k$, can be interpreted as some measure of the importance of intervals in terms of the presence of the global minimum point.
-The interval selection scheme for the next trial becomes clear -- the point of each subsequent trial $x^{k+1}$ from (\ref{eq:22}) is chosen in the interval with the maximum value of the interval characteristic (i.e., in the interval  where the global minimum is most likely to be located).
+   To explain the computation scheme under consideration, note the following.
+The calculated characteristics $R(i)$, $1<i \leq k$, can be interpreted as some measure of the importance of the intervals in terms of the presence of the global minimum point.
+The interval selection scheme for the next trial becomes obvious: the point $x^{k+1}$ of each subsequent trial from (\ref{eq:22}) is chosen in the interval with the maximum value of the interval characteristic (i.e., in the interval where the global minimum is most likely to be located).
 
-It should also be noted that the AGMIS computational scheme considered above refines the general scheme of global search for mixed-integer optimization problems from subsection \ref{subsec:31}.
+It should also be noted that the AGMIS computational scheme, considered above, refines the general scheme of global search for mixed-integer optimization problems from subsection \ref{subsec:31}.
 Thus, the choice of the function $F(\alpha, y(x), w_i)$, $1\leq i\leq l$, for the next iteration is provided by the procedure of finding the interval with the maximum characteristic.
@@
 
-An example of using the AGMIS algorithm for the problem (\ref{eq:12}) within the framework of the proposed approach is shown in Fig.~\ref{fig:1}.
-In this example, the function value $F(\lambda, y, u)$ for the discrete parameter value $u=1$ was calculated only 7 times, and for the value $u=2$, 17 times, i.e.
-the global search iteration was performed mainly for the value of the discrete parameter, which allows the smallest value of the function $F(\lambda, y, u)$ to be achieved.
+Figure~\ref{fig:1} depicts the use of AGMIS for the problem (\ref{eq:12}) under the proposed approach.
+In this example, the function value $F(\lambda, y, u)$ is calculated only seven times for the discrete-parameter value $u=1$, and seventeen times for the value $u=2$, that is, the global search iteration is performed mainly for the value of the discrete parameter, which makes it possible to achieve the smallest value of the function $F(\lambda, y, u)$.
 
-Now, let us return to the initial problem statement (\ref{eq:1}) and remind that for solving the MCOmix problem it may be necessary to solve the set of problems $\mathbb{F}_T$ from (\ref{eq:10}).
-In solving this set of problems by using the AGMIS algorithm, one more key property of the proposed approach is manifested -- the results of all previous  calculations of criteria values can be recalculated into the values of the next optimization problem $F(\alpha, y(x), u)$ from (\ref{eq:6}) to be solved with new values of $\alpha'$ from (\ref{eq:8}) without repeating any time-consuming calculations of criteria values, i.e.
+Now, let us return to the initial problem statement (\ref{eq:1}) and remember that for solving the MCOmix problem it may be necessary to solve the set of problems $\mathbb{F}_T$ from (\ref{eq:10}).
+One more key property of the proposed approach becomes apparent when solving this set of problems by the AGMIS algorithm: the results of all previous  computations of criteria values can be reused to calculate the values of the next optimization problem $F(\alpha, y(x), u)$ from (\ref{eq:6}), which is to be solved with new values of $\alpha'$ from (\ref{eq:8}), without having to repeat any time-consuming calculation of criteria values, that is,
 \begin{equation}\label{eq:24}
- z_i'=F(\alpha',y(x_i), u), 1 \leq i \leq k.
+ z_i'=F(\alpha',y(x_i), u),\ 1 \leq i \leq k.
 \end{equation}
 
-Thus, all search information $A_k$ from (\ref{eq:17}), recalculated according to (\ref{eq:24}), can be reused to continue solving the next problem $F(\alpha, y(x), u)$.
-This reuse of search information can provide a significant reduction in the amount of computations performed for each subsequent problem of the set $\mathbb{F}_T$ from (\ref{eq:10})  that may only require a relatively small number of global search iterations.
-This  has been confirmed  by our computational experiments -- see Section \ref{sec:5}.
+Thus, all search information $A_k$ from (\ref{eq:17}), recalculated according to (\ref{eq:24}), can be reused to continue the solution of the next problem $F(\alpha, y(x), u)$.
+This reuse of search information can significantly reduce the number of computations performed for each subsequent problem of the set $\mathbb{F}_T$ from (\ref{eq:10}), which may require only a relatively small number of global search iterations.
+This  has been confirmed  in our computational experiments (see Section \ref{sec:5}).
 
-The AGMIS algorithm, supplemented with the ability to reuse search information in MCOmix problems, will be further referred to as the Algorithm for Multicriteria Mixed-Integer Search (AMMIS).
+The AGMIS algorithm supplemented with the ability to reuse search information in MCOmix problems will be further referred to as the Algorithm for Multicriteria Mixed-Integer Search (AMMIS).
 
@@
 Unfortunately, attempts to develop parallel variants of AGMIS and AMMIS algorithms using existing parallelization methods have not succeeded.
-For example, data parallelization (splitting the computation domain between available computing elements) results in only one processor processing a subdomain of the search domain  containing the sought globally optimal solution of the problem, with other processors performing redundant calculations.
-In \cite{c24,c32} a new approach to parallelizing calculations for solving global optimization problems is proposed: parallelism of calculations is provided by organizing simultaneous calculation of the values of the minimized function $F(\alpha,y,u)$ from (\ref{eq:8}) at several different points of the search domain $D$.
-This approach makes it possible to parallelize the most time-consuming part of the global search process, and, due to its general nature -- it can be applied to almost any global search method for a wide variety of global optimization problems.
+For example, data parallelization (splitting the computation domain between available computing elements) results in only one processor processing the subdomain of the search domain containing the sought global optimal solution of the problem, while other processors perform redundant calculations.
+A new approach to parallelizing calculations for solving global optimization problems is proposed in \cite{c24,c32}: the parallelism of calculations is ensured by organizing simultaneous computation of values of the minimized function $F(\alpha,y,u)$ from (\ref{eq:8}) at several different points of the search domain~$D$.
+This approach makes it possible to parallelize the most time-consuming part of the global search process and, due to its general nature, it can be applied to almost any global search method for a wide variety of global optimization problems.
 
-Applying this approach and taking into account the interpretation of the characteristics $R(i)$, $1<i \leq k$, of search intervals $(x_{i-1}, x_i)$, $1<i \leq k$, from (\ref{eq:21}) as a measure of the intervals importance in terms of containing the global minimum point, a parallel version of the AGMIS algorithm can be obtained with the following generalization of rules (\ref{eq:21})-(\ref{eq:22}) \cite{c32,c40}:
+Applying this approach and taking into account the interpretation of the characteristics $R(i)$, $1<i \leq k$, of search intervals $(x_{i-1}, x_i)$, $1<i \leq k$, from (\ref{eq:21}) as a measure of the interval importance in terms of containing the global minimum point, we can obtain a parallel version of the AGMIS algorithm with the following generalization of rules (\ref{eq:21})--(\ref{eq:22}) (see \cite{c32,c40}):
 
-\textit{Rule 2 $'$}. Arrange the characteristics of intervals in descending order
+\textit{Rule 2$'$}. Arrange the characteristics of intervals in descending order,
 \begin{equation}\label{eq:25}
- R(t_1) \geq R(t_2) \geq \dots \geq R(t_{k-2}) \geq R(t_{k-1})
+ R(t_1) \geq R(t_2) \geq \dots \geq R(t_{k-2}) \geq R(t_{k-1}),
 \end{equation}
@@
 
-\textit{Rule 3 $'$}.
-Perform new trials (calculate the values of the minimized function $F(\alpha,y(x),u)$) at the points $x^{k+j}$, $1 \leq j \leq p$ located in the intervals with the maximum characteristics from (\ref{eq:25}).
+\textit{Rule 3$'$}.
+Perform new trials (calculate the values of the minimized function $F(\alpha,y(x),u)$) at the points $x^{k+j}$, $1 \leq j \leq p$, located in the intervals with the maximum characteristics from (\ref{eq:25}).
 
-The stopping condition for the algorithm (\ref{eq:23}) must be checked for all the intervals in which the next trials are performed
+Stopping condition (\ref{eq:23}) for the algorithm must be checked for all the intervals in which the next trials are performed,
 \begin{equation}\label{eq:26}
-(x_{t_j} - x_{t_j-1})^{1/n} \leq \varepsilon, 1 \leq j \leq p. 
+(x_{t_j} - x_{t_j-1})^{1/n} \leq \varepsilon,\ 1 \leq j \leq p.
 \end{equation}
 
-As in previous cases, if the stopping condition is not fulfilled, the number of iteration $k$ is increased by $p$, and a new global search iteration is executed.
+As in the cases considered before, if the stopping condition is not fulfilled, the number of iterations $k$ is increased by $p$, and a new global search iteration is executed.
 
-This parallel variant of the AGMIS algorithm will be further referred to as Parallel Algorthm for  Global Mixed-Integer Search (PAGMIS), and the parallel variant of the AMMIS algorithm will be referred to as PAMMIS, respectively.
+This parallel variant of the AGMIS algorithm will be further referred to as the Parallel Algorithm for Global Mixed-Integer Search (PAGMIS), and the parallel variant of the AMMIS algorithm, as PAMMIS.
 
-To assess the efficiency of the parallel algorithms, a large series of computational experiments has been carried out confirming that the proposed approach allows for a significant reducion in the amount of computation and time when solving complex multicriteria mixed-integer optimization problems -- see Section \ref{sec:5}.
+To assess the efficiency of the parallel algorithms, we conducted a large series of computational experiments, in which it was confirmed that the suggested approach can significantly reduce the number of computations and the time required to solve complex multicriteria mixed-integer optimization problems (see Section \ref{sec:5}).
 
-\section{Results of numerical experiments}\label{sec:5}
+\section{Results of Numerical Experiments}\label{sec:5}
 
-Numerical experiments were performed on the supercomputers Lobachevsky (University of Nizhni Novgorod), Lomonosov (Moscow State University), MVS-10P (Joint Supercomputer Center of RAS) and supercomputers Endeavor.
-The numerical results were obtained by using the following computational nodes: 2 Intel Xeon Platinum 8260L, 2.4 GHz, 256 GB RAM (i.e.
-a total of 48 CPU cores were available on each node).
-The executable program code was built by using the Intel Parallel Studio XE 2019 software package.
-The numerical experiments were performed using the Globalizer system \cite{c41}.
+The numerical experiments were performed on the supercomputers Lobachevsky (University of Nizhni Novgorod), Lomonosov (Moscow State University), MVS-10P (Joint Supercomputer Center of RAS), and Endeavor.
+Each computational node was equipped with two processors Intel Xeon Platinum 8260L 2.4 GHz (i.e., a total of 48 CPU cores on each node) and 256 GB of RAM.
+The executable program code was built with the software package Intel Parallel Studio XE 2019.
+For the numerical experiments, we used the Globalizer system \cite{c41}.
 
-The multiextremal optimization algorithms used in the framework of the proposed approach have demonstrated their effectiveness in numerical experiments and have been widely used in solving practical global search problems -- see, for example, \cite{c20,c21,c22,c23}.
-Below, we present the results of numerical experiments obtained earlier that prove the effectiveness of this approach \cite{c21}.
-In the experiments, the following bi-criteria test problem proposed in \cite{c42} was solved:
+The multiextremal optimization algorithms used under the suggested approach have demonstrated their effectiveness in numerical experiments and have been widely used to solve practical global search problems (see, for example, \cite{c20,c21,c22,c23}).
+Below, we present the results of previous numerical experiments obtained that prove the effectiveness of this approach \cite{c21}.
+The following bi-criteria test problem, proposed in \cite{c42}, was solved in the experiments:
 \begin{equation}\label{eq:27}
-f_1(y)=(y_1-1) y_2^2+1,f_2 (y)=y_2, 0 \leq y_1, y_2 \leq 1.
+f_1(y)=(y_1-1) y_2^2+1,f_2 (y)=y_2,\ 0 \leq y_1, y_2 \leq 1.
 \end{equation}
 
-The solution of the MCO problem consisted in constructing a numerical approximation of the Pareto set.
-To assess the quality of approximation, the completeness and uniformity of the Pareto set coverage was compared using the following two indicators \cite{c42,c43}:
+The solution of the MCO problem consisted of the construction of a numerical approximation of the Pareto set.
+To assess the quality of the approximation, the completeness and uniformity of the Pareto set coverage were compared using the following two indicators \cite{c42,c43}:
 \begin{itemize}
 	\item The hypervolume index (HV). This indicator evaluates the approximation of the Pareto set in terms of completeness (a higher value corresponds to a more complete coverage of the Pareto domain).
+
   \item The distribution uniformity index (DU). This indicator evaluates the uniformity of the Pareto domain coverage (a lower value corresponds to a more uniform coverage of the Pareto domain).
@@
 
-Five multicriteria optimization algorithms were compared in this experiment: the Monte-Carlo (MC) method \cite{c42}, the genetic algorithm SEMO from the PISA library \cite{c42}, the Non-uniform coverage (NUC) method \cite{c42}, the bi-objective Lipschitz optimization (BLO) method \cite{c43}, and the AMMIS method developed by the authors within the framework of the proposed approach.
+Five multicriteria optimization algorithms were compared in this experiment: the Monte-Carlo (MC) method \cite{c42}, the genetic algorithm SEMO from the PISA library \cite{c42}, the non-uniform coverage (NUC) method \cite{c42}, the bi-objective Lipschitz optimization (BLO) method \cite{c43}, and the AMMIS method, developed by the authors under the proposed approach.
 
-For AMMIS, 50 global optimization problems with different values of convolution coefficients $\lambda\in \Lambda$ from (\ref{eq:8}) were solved.
-The accuracy of the method $\varepsilon=0.05$ and the reliability parameter $r=3.0$ from (\ref{eq:23}) were used when solving the series of problems.
-The results of the experiments are presented in Table \ref{tab:1}.
+Fifty global optimization problems were solved with AMMIS for different values of the convolution coefficients $\lambda\in \Lambda$ from (\ref{eq:8}), with accuracy $\varepsilon=0.05$ and reliability parameter $r=3.0$ from (\ref{eq:23}).
+The results of the experiments are given in Table~\ref{tab:1}.
 
@@
 Solution method     & MC    & SEMO  & NUC   & BLO   & \textbf{AMMIS} \\ \hline
-Number of method iterations  & 500   & 500   & 515   & 498   & \textbf{370}   \\
+Number of iterations of the method & 500   & 500   & 515   & 498   & \textbf{370}   \\
 \begin{tabular}[c]{@{}c@{}}Number of points found in the Pareto domain\end{tabular} & 67    & 104   & 29    & 68    & \textbf{100}    \\
@@
 
-As revealed in the experiments, the AMMIS algorithm has a noticeable advantage over the multicriteria optimization methods considered here even when solving relatively simple MCO problems.
+The experiments confirmed that AMMIS has a noticeable advantage over the multicriteria optimization methods considered here, even when solving relatively simple MCO problems.
 
-The results of computational experiments to solve 100 MCOmix problems of the set $\mathbb{F}_T$ from (10) are given below.
+The results of the computational experiments with 100 MCOmix problems of the set $\mathbb{F}_T$ from (\ref{eq:10}) are given below.
 Each problem in the set $\mathbb{F}_T$ contained two criteria.
-Each criterion was formed using the functions obtained by the GKLS generator \cite{c44}, where some parameters were discrete \cite{c45}.
-To generate MCOmix problems, 4 continuous parameters $y$ from (\ref{eq:2}) and 5 discrete parameters $u$ from (\ref{eq:3}) were used, each of them only having two values (i.e., $n=4$, $m=5$, $l=32$).
+Each criterion was defined through functions obtained by the GKLS generator \cite{c44}, with some of the parameters being discrete \cite{c45}.
+To generate the MCOmix problems, we considered four continuous parameters $y$ from (\ref{eq:2}) and five discrete parameters $u$ from (\ref{eq:3}), with only two values for each of them (i.e., $n=4$, $m=5$, $l=32$).
 
-To solve each MCOmix problem, 50 uniformly distributed values of convolution coefficients $\lambda \in \Lambda$ from (\ref{eq:8}) were used.
-To estimate the values of HV and DU efficiency indicators, an approximation of the Pareto domain was calculated for each MCOmix problem  using a uniform grid in the search domain $D\times U$ from (\ref{eq:2})-(\ref{eq:3}).
-The PAMMIS algorithm was used for the following parameter values: accuracy $\varepsilon=0.05$ from (\ref{eq:26}) and reliability $r=5.6$ (the $r$ parameter is used in PAMMIS to construct estimates of the H{\"o}lder constant $H$ from (\ref{eq:16})).
+Fifty uniformly distributed values of the convolution coefficients $\lambda \in \Lambda$ from (\ref{eq:8}) were used to solve each MCOmix problem.
+To estimate the values of the HV and DU efficiency indicators, we calculated an approximation of the Pareto domain for each MCOmix problem, using a uniform grid in the search domain $D\times U$ from (\ref{eq:2})--(\ref{eq:3}).
+The following parameter values were used for PAMMIS: accuracy $\varepsilon=0.05$ from (\ref{eq:26}) and reliability $r=5.6$ (the $r$ parameter is used in PAMMIS to construct estimates of the H{\"o}lder constant $H$ in (\ref{eq:16})).
 
-Table \ref{tab:2} gives the results of numerical experiments, averaged by the number of MCOmix problems solved.
+Table \ref{tab:2} contains the results of the numerical experiments, averaged by the number of MCOmix problems solved.
 The first column shows the number of cores used.
-The second column contains information on the average number of trials (computation of criteria values) performed during the global search.
+The second column contains information on the average number of trials (computations of criteria values) performed during the global search.
 The third column shows the resulting speedup (reduction in the number of trials) when using parallel computations.
-The last two columns contain the values of HV and DU indicators.
+The last two columns show the values of the HV and DU indicators.
 
@@
 \centering
-\caption{Efficiency of the PAMMIS parallel algorithm in solving MCOmix problems}
+\caption{Efficiency of PAMMIS in the solution of MCOmix problems}
 \label{tab:2}
@@
 Cores                  & Iterations                  & Speedup                 & DU           & &       & HV                      \\ \hline
-\multicolumn{7}{c}{\begin{tabular}[c]{@{}c@{}}The Pareto domain estimate obtained \\ by exhaustive search\end{tabular}}        \\ \hline
+\multicolumn{7}{c}{\begin{tabular}[c]{@{}c@{}} Pareto domain estimate obtained \\ by exhaustive search\end{tabular}}        \\ \hline
 48                     & 1706667                   & -                       & 20.6         & &       & 29784.9                \\ \hline
-\multicolumn{7}{c}{\begin{tabular}[c]{@{}c@{}}The Pareto    domain  estimate obtained \\ using the PAMMIS method\end{tabular}} \\ \hline
+\multicolumn{7}{c}{\begin{tabular}[c]{@{}c@{}} Pareto    domain  estimate obtained \\ using the PAMMIS method\end{tabular}} \\ \hline
 1                      & 86261.6                    & 1.0                     & 13.8         & &      & 30212.3                \\
@@
 
-The results of the computational experiments show that the PAMMIS algorithm is scalable as the  speedup of parallel computations increases almost linearly with the number of computing cores used.
-The obtained values of HV and DU indicators show that the Pareto set estimate using the PAMMIS algorithm is calculated with a greater accuracy and requires a significantly smaller number of trials compared to the results obtained using uniform grids in the search domain $D \times U$.
+The results of the experiments demonstrate that PAMMIS is scalable as the  speedup of parallel computations increases almost linearly with the number of computing cores used.
+The obtained values of the HV and DU indicators indicate that the Pareto set estimate obtained by the PAMMIS method is calculated with greater accuracy and requires a significantly smaller number of trials compared to the results obtained by using uniform grids in the search domain $D \times U$.
 
-\section{Conclusion}
+\section{Conclusions}
+In this paper, we proposed a new approach to solving computationally intensive optimization problems, where some varied parameters can only take on discrete values (MCOmix).
+It is assumed that the efficiency criteria may be multiextremal, and the computation of criterion values may involve a large number of calculations.
+Due to the high computational complexity of this class of problems, parallel solution methods that rely on the efficient use of high-performance computing systems need to be developed.
 
-In this paper, a new approach is proposed to solving computationally intensive optimization problems, where some varied parameters can only take on discrete values (MCOmix).
-It is also assumed that the efficiency criteria may be multiextremal, and the computing of criterion values may involve a large amount of calculations.
-Due to the high computational complexity of this class of problems, parallel solution methods need to be developed, relying on the efficient use of high-performance computing systems.
+Under the proposed approach, the solution of mixed-integer optimization problems is reduced to solving a family of global search optimization problems that use only continuous parameters.
+All problems of the family are solved simultaneously in time-shared mode, where the optimization problem for the next global search iteration is selected adaptively taking into account the search information obtained in the course of the calculations.
+The results of numerical experiments prove that  this approach can significantly reduce the computational intensity when solving MCOmix problems.
 
-In the framework of the proposed approach, solution of mixed-integer optimization problems is reduced to solving a family of global search optimization problems that use only continuous parameters.
-All problems of the family are solved simultaneously in the time-shared mode, where the optimization problem for the next global search iteration is selected adaptively taking into account the search information obtained in the course of computations.
-The results of numerical experiments prove that  this approach can significantly reduce computational intensity when solving MCOmix problems.
-
-In conclusion, it may be noted that the proposed approach is promising and requires continued investigations.
-First of all, it is necessary to continue numerical experiments to solve multicriteria mixed-integer optimization problems with a greater number of efficiency criteria and having a greater dimensionality.
-The possibility of organizing parallel computations using high-performance systems with distributed memory should be also evaluated.
+Finally, it should be noted that the approach we propose is quite promising and requires further research.
+First of all, it is necessary to perform numerical experiments involving multicriteria mixed-integer optimization problems with a greater number of efficiency criteria and higher dimensionality.
+It is also necessary to assess the possibility of implementing parallel computations on high-performance systems with distributed memory.
 
 \section*{Acknowledgements}
-This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.0729-2020-0055, and by the Research and Education Mathematical Center, project no.
-075-02-2020-1483/1.
+The work was supported by the Ministry of Science and Higher Education of the Russian Federation (project no.0729-2020-0055) and by the Research and Education Mathematical Center (project no.
+075-02-2020-1483/1).
 

### 2021-04-05 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
Правки MOTOR

@@
 
-Having several objective functions in MOO problems that determine the effectiveness of the decisions to be made results in most cases in situations where no option can be found in the whole set of alternatives that would be the best one with respect to all efficiency criteria at the same time.
-In such situations, the solution of MOO problems usually consists in finding non-dominated (or efficient) decisions in which values cannot be improved simultaneously across all objective functions.
-The set of all efficient decisions is usually referred to as the Pareto set.
-
-Finding the whole Pareto set when solving MOO problems can be computationally intensive (especially when objective functions are multiextremal).
-Moreover, having the whole Pareto set may be redundant for the decision maker (DM) due to the complexity of analyzing a large number of possible alternatives.
-As a result, when solving MOO problems in practice, it may be sufficient to find one or several efficient decisions taking into account the DM's perceptions of the required optimality of the options to be chosen.
-
 One of the approaches most commonly used in the search for efficient decisions is scalarization of the vector efficiency criterion, when a MOO problem is reduced to solving one or several scalar (in the general case, global) optimization problems in which the objective functions are converted to the integrated scalar function.
@@
 One common property of methods based on scalarization of the vector efficiency criterion is the existence of some scalarization coefficients that can be varied to obtain different solutions from the Pareto set.
-Thus, the scalarization coefficients may be interpreted as measures of the importance of the efficiency criteria determined by DM according to his/her perception of the required optimality of the decisions to be made.
+Thus, the scalarization coefficients may be interpreted as measures of the importance of the efficiency criteria determined by the decision maker (DM) according to his/her perception of the required optimality of the decisions to be made.
 As a result, the general scheme for solving the MOO problem can be represented as a sequence of steps; at each step, DM sets the necessary scalarization coefficients, then the resulting scalar optimization problem is solved, after that DM analyzes the efficient decision found and, if necessary, the above steps are repeated.
@@
 
-In the most general form, the multi-objective optimization (MOO) problem can be formulated as follows
+In the most general form, the MCO problem can be formulated as follows
 \begin{equation}
@@
 where $f(y) = (f_1(y), f_2(y), \dots, f_s(y))$ are objective functions (efficiency criteria), $y = (y_1, y_2, \dots , y_N)$ is the vector of varied parameters, and $N$ is the dimensionality of the multi-objective optimization problem to be solved.
-The set of possible parameter values (search domain) $D$ is usually an $N$-dimensional hypercube
+The set of possible parameter values (search domain) $D$ is usually an $N$-dimensional hyperinterval
 \begin{equation}
@@
 where $L_i$, $1 \leq i \leq s$, are the Lipschitz constants and $\|*\|$ denotes the Euclidean norm in $R^N$.
-Condition (\ref{eq:03}) means that for small variations of the parameter $y \in D$, the corresponding changes in the values of the functions $f_i(y)$, $1 \leq i \leq s$, are bounded.
+Condition (\ref{eq:03}) means that for variations of the parameter $y \in D$, the corresponding changes in the values of the functions $f_i(y)$, $1 \leq i \leq s$, are bounded.
 
@@
 
-\textbf{1.
-Scalarization of multiple objective functions.} In the most general form, the global optimization problem generated by the scalarization of multiple objective functions of the MOO problem can be represented in the form
+\subsection{Scalarization of multiple objective functions.} 
+In the most general form, the global optimization problem generated by the scalarization of multiple objective functions of the MOO problem can be represented in the form
 \begin{equation}
@@
 
-The reference decision $y^0 \in D$ in (\ref{eq:05}) can be known \textit{a priori} or determined on the basis of some known prototype.
+The reference decision $y^0 \in D$ in (\ref{eq:05}) can be known \textit{a priori} or determined on a basis of some known prototype.
 In many cases, an abstract ideal decision $y^0 \in D$ is used as the reference decision $y^*\in D$ in which the objective functions $f_i$, $1 \leq i \leq s$, have minimal possible values, i.e.
@@
 
-\textbf{2.
-Dimensionality reduction.} As already mentioned, problem (\ref{eq:04}) is a multidimensional global optimization problem.
+\subsection{Dimensionality reduction.} 
+As already mentioned, problem (\ref{eq:04}) is a multidimensional global optimization problem.
 Problems of this kind are computationally complex and are known to be subject to the ``curse of dimensionality'' -- computational complexity increases exponentially with increasing dimensionality of the optimization problem being solved \cite{c13,c14,c15,c16,c17,c18,c19,c20}.
@@
 
-\textbf{3.
-Solving the one-dimensional reduced optimization problem.} Using the dimensionality reduction results in one more additional advantage of the proposed approach: many well-known one-dimensional global search algorithms (possibly, after some additional generalisation) can be used to solve the initial multidimensional MOO problem from (\ref{eq:01}) \cite{c22,c23,c24,c25,c26,c27,c28,c29}.
+\subsection{Solving the one-dimensional reduced optimization problem.} 
+Using the dimensionality reduction results in one more additional advantage of the proposed approach: many well-known one-dimensional global search algorithms (possibly, after some additional generalisation) can be used to solve the initial multidimensional MOO problem from (\ref{eq:01}) \cite{c22,c23,c24,c25,c26,c27,c28,c29}.
 At the same time, it should be noted that most works where dimensionality reduction is a key feature for solving multiextremal problems like (\ref{eq:04}) rely on the information-statistical theory of global search \cite{c15}.

### 2021-04-07 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
MOTOR правки

@@
 In such cases, all the search information obtained in solving any single scalar problem can be used for solving all other scalar problems of the same group.
+
+This paper presents the results of the authors continuing research aimed to develop effective methods for solving MOO problems \cite{c35,c36,c37,c44}.
+In \cite{c35}, some methods for solving global optimization problems are considered, in \cite{c36},  a general decision-making model based on MOO problems is proposed, and in \cite{c37}, a theoretical analysis is presented regarding the effectiveness of the use of search information when solving MOO problems.
+In \cite{c44}, we consider MOO problems in which the efficiency criteria can be ordered by importance (lexicographic multi-objective optimization).
+In this case, MOO problems are reduced to global optimization problems with non-convex constraints for which constrained optimization methods should be applied.
+In this paper, we propose new computational schemes for using search information that significantly reduce the amount of calculations required to solve MOO problems.
 
@@
 
-\textbf{1.
-The need to solve multiple global optimization problems.} As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different efficient decisions due to possible changes in the optimality requirements.
+In this Section, the proposed approach for finding $p$, $p>1$ Pareto-optimal decisions of the MOO problem is presented.
+This problem is formulated as the problem of solving the set $\Phi_p (y)$, $p>1$ of global optimization problems (see \ref{subsec:41}).
+To solve this set of optimization problems, two efficient computational schemes are proposed.
+The first scheme is based on the traditional sequential procedure for solving multiple problems, however, to solve each subsequent optimization problem, all the search information obtained during previous calculations is taken into account.
+This accumulated search information can significantly reduce the number of global search iterations required to solve the next optimization problems of the set $\Phi_p (y)$ (see \ref{subsec:42} and Chapter \ref{sec:05}).
+In the second scheme, all the problems of the set $\Phi_p (y)$ are solved simultaneously, but the search information obtained when solving each problem is also used for solving all the other problems of the set $\Phi_p (y)$ (see \ref{subsec:43}).
+This useful exchange of search information also significantly reduces the amount of required calculations for solving the problems of the set $\Phi_p (y)$.
+
+\subsection{The need to solve multiple global optimization problems.} 
+\label{subsec:41}
+
+As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different efficient decisions due to possible changes in the optimality requirements.
 Obtaining different efficient decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multi-extremal functions $F(\lambda, y)$ from (\ref{eq:04}).
@@
 
-\textbf{2.
-Step-by-step solution of a set of global optimization problems.} As shown in Section \ref{sec:03}, the information-statistical multiextremal optimization algorithms used in the proposed approach determine the points of consecutive iterations of the global search taking into account the search information
+\subsection{Step-by-step solution of a set of global optimization problems.} 
+\label{subsec:42}
+
+As shown in Section \ref{sec:03}, the information-statistical multiextremal optimization algorithms used in the proposed approach determine the points of consecutive iterations of the global search taking into account the search information
 \begin{equation}
@@
 
-\textbf{3.
-Simultaneous solution of a set of global optimization problems.} Information connectivity makes it possible to propose a more general scheme for simultaneous optimization of all the functions of the set $\Phi_p(y)$ from (\ref{eq:21}).
+\subsection{Simultaneous solution of a set of global optimization problems.} 
+\label{subsec:43}
+
+Information connectivity makes it possible to propose a more general scheme for simultaneous optimization of all the functions of the set $\Phi_p(y)$ from (\ref{eq:21}).
 In this case, the search information $A_k$ from (\ref{eq:22}) will contain the computed values of all simultaneously optimized functions $F(\lambda_i, y)$, $1 \leq i \leq p$, i.e.
@@
 
-The results of numerical experiments are presented in Table \ref{tab:2}.
+The results of numerical experiments are presented in Table \ref{tab:2} and fig.
+\ref{fig:1}.
 The table shows in the first column the number of functions $\Phi_p (y)$ from (\ref{eq:21}) being optimized.
@@
 \end{table}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.7\linewidth]{fig_t2}
+  \caption{The average number of iterations for solving bi-criteria two-dimensional MOO problems}
+  \label{fig:1}
+\end{figure}
 
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.7\linewidth]{fig_t5}
+  \caption{The average number of iterations for solving bi-criteria five-dimensional MOO problems}
+  \label{fig:1}
+\end{figure}
+
 The results of our experiments show that with increasing dimensionality of the MOO problems to be solved, the trend that has been identified continues: the amount of computation (the number of global search iterations) is reduced by more than 5.6 times due to reuse of search information, the MGSA algorithm performs the least number of iterations, and the SGSA algorithm provides the best approximation of the Pareto domain.
@@
 \doi{10.1016/j.ejor.2009.12.027}
+\bibitem{c35} Gergel, V.: An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems.
+2nd International Conference on Mathematics and Computers in Sciences and in Industry, 13--18 (2015).
+\doi{10.1109/MCSI.2015.18}
+\bibitem{c36} Gergel, V.P., Kozinov, E.A.: Accelerating multicriterial optimization by the intensive exploitation of accumulated search data.
+AIP Conference Proceedings, \textbf{1776}, 090003 (2016).
+\doi{10.1063/1.4965367}
+\bibitem{c37} Gergel, V., Kozinov, E. Efficient multicriterial optimization based on intensive reuse of search information. Journal of Global Optimization 71(1), 73--90 (2018).
+\bibitem{c44} Gergel, V., Kozinov, E., Barkalov, K.: Computationally efficient approach for solving lexicographic multicriteria optimization problems.
+Optim.
+Lett.
+(2020).
+\doi{10.1007/s11590-020-01668-y}
 \bibitem{c13} Zhigljavsky, A.A.: Theory of Global Random Search. Dordrecht: Kluwer Academic Publishers (1991).
@@
 \bibitem{c34} Grishagin, V., Israfilov, R., Sergeyev, Y.: Comparative efficiency of dimensionality reduction schemes in global optimization. AIP Conference Proceedings 1776 (2016).
-\bibitem{c35} Gergel, V.: An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems.
-2nd International Conference on Mathematics and Computers in Sciences and in Industry, 13--18 (2015).
-\doi{10.1109/MCSI.2015.18}
-\bibitem{c36} Gergel, V.P., Kozinov, E.A.: Accelerating multicriterial optimization by the intensive exploitation of accumulated search data.
-AIP Conference Proceedings, \textbf{1776}, 090003 (2016).
-\doi{10.1063/1.4965367}
-\bibitem{c37} Gergel, V., Kozinov, E. Efficient multicriterial optimization based on intensive reuse of search information. Journal of Global Optimization 71(1), 73--90 (2018).
 \bibitem{c38} Gergel, V., Barkalov, K., Sysoyev, A.
@@
 ACM Transactions on Mathematical Software, \textbf{29}(4), 469--480 (2003).
+
 \end{thebibliography}

### 2021-04-08 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
MOTOR правки

@@
 %
-\title{An Approach for Simultaneous Finding of Multiple Effective Decisions in Multi-objective Optimization Problems
+\title{An Approach for Simultaneous Finding of Multiple Efficient Decisions in Multi-objective Optimization Problems
 %with non-convex constraints 
@@
 
-In the most general form, the MCO problem can be formulated as follows
+In the most general form, the MOO problem can be formulated as follows
 \begin{equation}
@@
 
-\subsection{Scalarization of multiple objective functions.} 
+\subsection{Scalarization of multiple objective functions} 
 In the most general form, the global optimization problem generated by the scalarization of multiple objective functions of the MOO problem can be represented in the form
@@
 
-\subsection{Dimensionality reduction.} 
+\subsection{Dimensionality reduction} 
 As already mentioned, problem (\ref{eq:04}) is a multidimensional global optimization problem.
@@
 
-\subsection{Solving the one-dimensional reduced optimization problem.} 
+\subsection{Solving the one-dimensional reduced optimization problem} 
 Using the dimensionality reduction results in one more additional advantage of the proposed approach: many well-known one-dimensional global search algorithms (possibly, after some additional generalisation) can be used to solve the initial multidimensional MOO problem from (\ref{eq:01}) \cite{c22,c23,c24,c25,c26,c27,c28,c29}.
@@
 The first scheme is based on the traditional sequential procedure for solving multiple problems, however, to solve each subsequent optimization problem, all the search information obtained during previous calculations is taken into account.
-This accumulated search information can significantly reduce the number of global search iterations required to solve the next optimization problems of the set $\Phi_p (y)$ (see \ref{subsec:42} and Chapter \ref{sec:05}).
+This accumulated search information can significantly reduce the number of global search iterations required to solve the next optimization problems of the set $\Phi_p (y)$ (see \ref{subsec:42} and Section \ref{sec:05}).
 In the second scheme, all the problems of the set $\Phi_p (y)$ are solved simultaneously, but the search information obtained when solving each problem is also used for solving all the other problems of the set $\Phi_p (y)$ (see \ref{subsec:43}).
@@
 
-\subsection{The need to solve multiple global optimization problems.} 
+\subsection{The need to solve multiple global optimization problems} 
 \label{subsec:41}
@@
 
-\subsection{Step-by-step solution of a set of global optimization problems.} 
+\subsection{Step-by-step solution of a set of global optimization problems} 
 \label{subsec:42}
@@
 
-\subsection{Simultaneous solution of a set of global optimization problems.} 
+\subsection{Simultaneous solution of a set of global optimization problems} 
 \label{subsec:43}
@@
 When solving the problems, the accuracy $\varepsilon=0.05$ and the reliability $r=5.6$ were used.
-The results of numerical experiments are presented in Table \ref{tab:4}.
+The results of numerical experiments are presented in Table \ref{tab:4} and fig.
+\ref{fig:2}.
 
@@
   \caption{The average number of iterations for solving bi-criteria five-dimensional MOO problems}
-  \label{fig:1}
+  \label{fig:2}
 \end{figure}

### 2021-04-08 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
MOTOR правки

@@
 %
-\titlerunning{Finding of Multiple Effective Decisions in MOO Problems}
+\titlerunning{Finding of Multiple Efficient Decisions in MOO Problems}
 % If the paper title is too long for the running head, you can set
@@
 
-Choosing the optimal options (decision making) in situations with many different alternatives is a problem that occurs in almost every domain of human activity.
+Choosing the optimal decisions (decision making) in situations with many different alternatives is a problem that occurs in almost every domain of human activity.
 In many cases, decision making problems can be viewed as optimization problems of various kinds such as convex programming, discrete optimization, nonlinear programming, etc.
@@
 This effect is achived due to the fact that all such scalar problems  are based on the same MOO problem and, consequently, all computed values of the efficiency criteria of the MOO problem can be reduced to values of any scalar problem being solved simultaneously without any time consuming calculations.
-In such cases, all the search information obtained in solving any single scalar problem can be used for solving all other scalar problems of the same group.
+In such cases, all the search information obtained in solving any single scalar problem can be used for solving all other scalar problems of the same set.
 
-This paper presents the results of the authors continuing research aimed to develop effective methods for solving MOO problems \cite{c35,c36,c37,c44}.
+This paper presents the results of the authors continuing research aimed to develop efficient methods for solving MOO problems \cite{c35,c36,c37,c44}.
 In \cite{c35}, some methods for solving global optimization problems are considered, in \cite{c36},  a general decision-making model based on MOO problems is proposed, and in \cite{c37}, a theoretical analysis is presented regarding the effectiveness of the use of search information when solving MOO problems.
@@
 Section \ref{sec:05} contains the results of numerical experiments confirming the effectiveness of the proposed approach.
-In conclusion, the results obtained are discussed and possible main lines of action for further research are outlined.
+In conclusion, the results obtained are discussed and possible main directions of further research are outlined.
 
@@
 \subsection{Scalarization of multiple objective functions}
+\label{subsec:31}
+
 In the most general form, the global optimization problem generated by the scalarization of multiple objective functions of the MOO problem can be represented in the form
@@
 \end{equation}
-where $F(\lambda,y)$ is the standard deviation of the values of the objective functions $f_i$, $1 \leq i \leq s$, for the decision $y \in D$ and for the specified reference decision $y^0 \in D$, while the coefficients $\lambda_i$, $0 \leq \lambda_i \leq 1$, $1 \leq i \leq s$, are measures of the significance of the approximation accuracy for each variable $y_i$, $1 \leq i \leq N$, separately.
-Without loss of generality, we can assume that the area of possible values of the coefficients $\lambda$ is a set
+where $F(\lambda,y)$ is the standard deviation of the values of the objective functions $f_i$, $1 \leq i \leq s$, for the decision $y \in D$ and for the specified reference decision $y^0 \in D$, while the coefficients $\lambda_i$, $0 \leq \lambda_i \leq 1$, $1 \leq i \leq s$, are measures of the importance of the approximation accuracy for each variable $y_i$, $1 \leq i \leq N$, separately.
+Without loss of generality, we can assume that the domain of possible values of the coefficients $\lambda$ is a set
 \begin{equation}
@@
 \subsection{Dimensionality reduction} 
+\label{subsec:32}
+
 As already mentioned, problem (\ref{eq:04}) is a multidimensional global optimization problem.
 Problems of this kind are computationally complex and are known to be subject to the ``curse of dimensionality'' -- computational complexity increases exponentially with increasing dimensionality of the optimization problem being solved \cite{c13,c14,c15,c16,c17,c18,c19,c20}.
-Nevertheless, the computational complexity of global optimization algorithms can be significantly reduced by dimensionality reduction based on the use of a Peano space-filling curve (or evolvent) $y(x)$ that uniquely and continuously maps the segment $[0,1]$ on an $N$-dimensional hypercube $D$ -- see, for example, \cite{c15,c21}.
+Nevertheless, the computational complexity of global optimization algorithms can be significantly reduced by dimensionality reduction based on the use of a Peano space-filling curve (or evolvent) $y(x)$ that uniquely and continuously maps the segment $[0,1]$ on an $N$-dimensional domain $D$ -- see, for example, \cite{c15,c21}.
 As a result of such reduction, multidimensional global optimization problems (\ref{eq:04}) are reduced to one-dimensional problems
@@
 \end{equation}
-where the constant $H$ is defined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:08}), and $N$ is the dimensionality of the optimization problem (\ref{eq:01}).
+where the constant $H$ is defined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:08}), and $N$ is the dimensionality of the MOO problem (\ref{eq:01}).
 
 \subsection{Solving the one-dimensional reduced optimization problem} 
+\label{subsec:33}
+
 Using the dimensionality reduction results in one more additional advantage of the proposed approach: many well-known one-dimensional global search algorithms (possibly, after some additional generalisation) can be used to solve the initial multidimensional MOO problem from (\ref{eq:01}) \cite{c22,c23,c24,c25,c26,c27,c28,c29}.
 At the same time, it should be noted that most works where dimensionality reduction is a key feature for solving multiextremal problems like (\ref{eq:04}) rely on the information-statistical theory of global search \cite{c15}.
-This theory has provided the a basis for developing a large number of efficient methods for multi-extremal optimization \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
+This theory has provided the a basis for developing a large number of efficient methods for multiextremal optimization \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
 
@@
 \label{eq:13}
-R(i) = 0.5 H \rho_i - 0.5 (z_{i-1} + z_i ), \rho_i = (x_i-x_{i-1} ) ^ {1⁄N}, 1 < i \leq k,
+R(i) = 0.5 H \rho_i - 0.5 (z_{i-1} + z_i ), \rho_i = (x_i-x_{i-1} )^{1⁄N}, 1 < i \leq k,
 \end{equation}
@@
 \label{eq:14}
-R(i)=m\rho_i+\frac{z_i-z_{i-1}^2}{m\rho_i}-2(z_{i-1}+z_i ),\rho_i=(x_i-x_{i-1})^{1/N}, 1 < i \leq k,
+R(i)=m\rho_i+\frac{(z_i-z_{i-1})^2}{m\rho_i}-2(z_{i-1}+z_i ),\rho_i=(x_i-x_{i-1})^{1/N}, 1 < i \leq k,
 \end{equation}
@@
 
-The presence of interval characteristics makes it possible to present the procedure of global search iteration as the following sequence of steps \cite{c15}.
+The presence of interval characteristics makes it possible to describe the procedure of global search iteration as the following sequence of steps \cite{c15}.
 
@@
 \label{eq:19}
-z_k^* = \min \{ z_i, 1 \leq i \leq \}.
+z_k^* = \min \{ z_i, 1 \leq i \leq k\}.
 \end{equation}
@@
 
-Convergence conditions of the algorithms formulated in the form (\ref{eq:16})--(\ref{eq:18}).
+Convergence conditions of the algorithms formulated in the form (\ref{eq:16})--(\ref{eq:18}) depend on the properties of the interval characteristics used.
 One of sufficient conditions for convergence of algorithms is, for example, the requirement that the characteristic of the interval containing the global minimum point should take on the maximum value at step 1 of the scheme (\ref{eq:16})--(\ref{eq:18}) during the global search iterations.
 This condition is satisfied, for example, for the multidimensional generalized algorithms proposed in \cite{c22,c23} when the H{\"o}lder constant from (\ref{eq:10}) is specified exactly.
-For GSA, a sufficient condition for convergence is the relation \cite{c15}
+For GSA, a sufficient condition for convergence is the relation
 \begin{equation}
@@
 \end{equation}
-which must be fulfilled starting from some iteration $k>1$ of the global search ($L$ is the Lipschitz constant from (\ref{eq:08})).
+which must be fulfilled starting from some iteration $k>1$ of the global search ($L$ is the Lipschitz constant from (\ref{eq:08})) \cite{c15}.
 Moreover, if condition (\ref{eq:20}) is satisfied, only the points of the global minimum of the function $\varphi(y)$ from (\ref{eq:04}) will be the limit points of the trial sequence $\{y^k=y(x^k)\}$ generated by the GSA algorithm.
@@
 As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different efficient decisions due to possible changes in the optimality requirements.
-Obtaining different efficient decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multi-extremal functions $F(\lambda, y)$ from (\ref{eq:04}).
+Obtaining different efficient decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multiextremal functions $F(\lambda, y)$ from (\ref{eq:04}).
 These functions $F(\lambda, y)$ can be optimized sequentially.
@@
 
-The general scheme considered above can be extended by the option of selecting not one but several different scalarization coefficients at each stage.
+The general scheme considered above can be extended by the possibility of selecting not one but several different scalarization coefficients at each stage.
 Having such an option makes it easier for DM  to specify importance coefficients $\lambda \in \Lambda$ of the objective functions $f_i(y)$, $1 \leq i \leq s$.
@@
 
-Such generalization of the process of solving a MOO problem means that at each current moment of calculations there is a set of functions being optimized simultaneously having the form (\ref{eq:04})
+Such generalization of the process of solving a MOO problem means that at each current moment of calculations there is a set of functions being optimized simultaneously having the form
 \begin{equation}
@@
 obtained in the calculations (see (\ref{eq:14})--(\ref{eq:17})).
-In (\ref{eq:22}) $x_i$, $1 \leq i \leq k$,  are the reduced points of performed global search iterations ordered in ascending order of coordinates,  $z_i$, $f_i$, $1 \leq i \leq k$, are values of the scalar function $F(\lambda, y)$, from (\ref{eq:04}) and objective functions $f(y)$ from (\ref{eq:01}) of the current optimization problem to be solved at the points $x_i=y(x_i)$, $1 \leq i \leq k$.
+In (\ref{eq:22}) $x_i$, $1 \leq i \leq k$,  are the reduced points of performed global search iterations ordered in ascending order of coordinates,  $z_i$, $f_i$, $1 \leq i \leq k$, are values of the scalar function $F(\lambda, y)$, from (\ref{eq:04}) and objective functions $f_i(y)$, $1 \leq i \leq s$, from (\ref{eq:01}) of the current optimization problem to be solved at the points $x_i=y(x_i)$, $1 \leq i \leq k$.
 By using search information $A_k$ from (\ref{eq:22}) when choosing next search iterations it is possible to solve global optimization problems more efficiently and to provide convergence of algorithms only to the global minima of multiextremal functions being minimized.
 
-It is important to note that since the set $\Phi_p(y)$ of functions being optimized  simultaneously is generated from the same MOO problem from (\ref{eq:01}), the existence of the set $A_k$ from (\ref{eq:22}) allows us to adjust the results of all previously performed calculations of the values of the objective functions $f(y)$ to the values of the next optimized function $F(\lambda, y)$ from (\ref{eq:04}) without repeating any time consuming calculations of values, i.e.
+It is important to note that since the set $\Phi_p(y)$ of functions being optimized  simultaneously is generated from the same MOO problem from (\ref{eq:01}), the existence of the set $A_k$ from (\ref{eq:22}) allows us to adjust the results of all previously performed calculations of the values of the objective functions $f_i(y)$, $1 \leq i \leq s$, to the values of the next optimized function $F(\lambda, y)$ from (\ref{eq:04}) without repeating any time consuming calculations of values, i.e.
 \begin{equation}
@@
 \label{eq:29}
-m_j = r M_j, M_j=\left\{ \frac{\max|z_i (j)-z_{i-1} (j)|}{\rho_i} ,1< i \leq k, 1 \leq j \leq \right\}.
+m_j = r M_j, M_j=\left\{ \frac{\max|z_i (j)-z_{i-1} (j)|}{\rho_i} ,1< i \leq k, 1 \leq j \leq p \right\}.
 \end{equation}
@@
 
-Numerical experiments were performed on the Lobachevsky supercomputer of the University of Nizhni Novgorod (operating system -- CentOS 6.4, control system -- SLURM).
+Numerical experiments were performed on the Lobachevsky supercomputer of the University of Nizhni Novgorod (operating system -- CentOS 6.4, management system -- SLURM).
 One supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM.
@@
 For MGSA 100 problems (\ref{eq:04}) with different values of convolution coefficients $\lambda$ uniformly distributed in $\Lambda$ were solved.
-The full results of the experiments performed are presented in Table \ref{tab:1}.
+The results of the experiments performed are presented in Table \ref{tab:1}.
 
@@
 
-The results of numerical experiments are presented in Table \ref{tab:2} and fig.
+The results of numerical experiments are presented in Table \ref{tab:2} and Fig.
 \ref{fig:1}.
-The table shows in the first column the number of functions $\Phi_p (y)$ from (\ref{eq:21}) being optimized.
+The Table shows in the first column the number of functions $\Phi_p (y)$ from (\ref{eq:21}) being optimized.
 The columns named Iters contain the average number of iterations performed by the algorithm to solve the MOO problem.
 The columns named HV and DU contain the values of the HV and DU indicators.
-The results of the numerical experiments are divided with respect to the three algorithms used: GSA is described in Section \ref{sec:03}, MGSA is discussed in Section \ref{sec:03} (item 2), SGSA is presented in Section \ref{sec:04} (item 3).
+The results of the numerical experiments are divided with respect to the three algorithms used: GSA is described in Section \ref{sec:03}, MGSA is discussed in Section \ref{subsec:32}, SGSA is presented in Section \ref{subsec:43}.
 
@@
 In the third series of numerical experiments, 10 bi-criteria five-dimensional MOO problems were solved, i.e.
-$N = 5$, $s = 2$, $m=2$.
+$N = 5$, $s = 2$.
 The objective functions of the MOO problems were determined, as before, using the GKLS generator \cite{c43}.
 When solving the problems, the accuracy $\varepsilon=0.05$ and the reliability $r=5.6$ were used.
-The results of numerical experiments are presented in Table \ref{tab:4} and fig.
-\ref{fig:2}.
+The results of numerical experiments are presented in Table \ref{tab:4} and Fig.\ref{fig:2}.
 
@@
 
-This paper proposes an efficient method for solving computationally intensive multi-objective optimization problems which require computing several Pareto-optimal decisions.
+This paper proposes an efficient approach for solving computationally intensive multi-objective optimization problems which require computing several Pareto-optimal decisions.
 It is also assumed that objective functions may be multiextremal, and computation of function values may require a large amount of calculations.
@@
 \bibitem{c3} Collette, Y., Siarry, P.:  Multiobjective Optimization: Principles and Case Studies (Decision Engineering). Springer (2011).
-\bibitem{c4} Marler, R. T., Arora, J. S.: Multi-Objective Optimization: Concepts and Methods for Engineering. VDM Verlag (2009).
+\bibitem{c4} Marler, R.T., Arora, J.S.: Multi-Objective Optimization: Concepts and Methods for Engineering. VDM Verlag (2009).
 \bibitem{c5} Pardalos, P.M., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-Convex Multi-Objective Optimization. Springer (2017).
-\bibitem{c6} Marler, R. T., Arora, J. S. Survey of multi-objective optimization methods for engineering. Struct. Multidisciplinary Optimization \textbf{26}, 369--395 (2004).
-\bibitem{c7}  Figueira, J., Greco, S., Ehrgott, M., eds.: Multiple criteria decision analysis: State of the art surveys. New York (NY). Springer (2005).
-\bibitem{c8} Zavadskas, E. K., Turskis, Z., Kildien{\. e}, S.: State of art surveys of overviews on MCDM/MADM methods. Technological and Economic Development of Economy, \textbf{20}, 165--179 (2014).
+\bibitem{c6} Marler, R.T., Arora, J.S.: Survey of multi-objective optimization methods for engineering. Struct. Multidisciplinary Optimization \textbf{26}, 369--395 (2004).
+\bibitem{c7}  Figueira, J., Greco, S., Ehrgott, M., eds.: Multiple criteria decision analysis: state of the art surveys. New York (NY). Springer (2005).
+\bibitem{c8} Zavadskas, E.K., Turskis, Z., Kildien{\. e}, S.: State of art surveys of overviews on MCDM/MADM methods. Technological and Economic Development of Economy, \textbf{20}, 165--179 (2014).
 \bibitem{c9} Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications. Surv. Math. Ind. \textbf{11}, 1--42 (2005).
@@
 \bibitem{c12} Figueira, J., Liefooghe, A., Talbi, E., Wierzbicki, A.: A parallel multiple reference point approach for multi-objective optimization.
-European Journal of Operational Research, \textbf{205}(2), 390--400 2010.
+European Journal of Operational Research \textbf{205}(2), 390--400 (2010).
 \doi{10.1016/j.ejor.2009.12.027}
-\bibitem{c35} Gergel, V.: An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems.
+\bibitem{c35} Gergel, V.: An unified approach to use of coprocessors of various types for solving global optimization problems.
 2nd International Conference on Mathematics and Computers in Sciences and in Industry, 13--18 (2015).
@@
 \bibitem{c36} Gergel, V.P., Kozinov, E.A.: Accelerating multicriterial optimization by the intensive exploitation of accumulated search data.
-AIP Conference Proceedings, \textbf{1776}, 090003 (2016).
+AIP Conference Proceedings \textbf{1776}, 090003 (2016).
 \doi{10.1063/1.4965367}
-\bibitem{c37} Gergel, V., Kozinov, E. Efficient multicriterial optimization based on intensive reuse of search information. Journal of Global Optimization 71(1), 73--90 (2018).
+\bibitem{c37} Gergel, V., Kozinov, E.: Efficient multicriterial optimization based on intensive reuse of search information. Journal of Global Optimization \textbf{71}(1), 73--90 (2018).
 \bibitem{c44} Gergel, V., Kozinov, E., Barkalov, K.: Computationally efficient approach for solving lexicographic multicriteria optimization problems.
@@
 \bibitem{c20} Sergeyev, Y.D., Kvasov, D.E.: Deterministic Global Optimization: An Introduction to the Diagonal Approach. Springer, New York (2017). 
-\bibitem{c21} Sergeyev Y.D., Strongin R.G., Lera D.: Introduction to global optimization exploiting space-filling curves. Springer (2013).
-\bibitem{c22} Piyavskij, S.: An algorithm for finding the absolute extremum of a function, Computational Mathematics and Mathematical Physics \textbf{12}, 57--67 (1972).
+\bibitem{c21} Sergeyev, Y.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013).
+\bibitem{c22} Piyavskij, S.: An algorithm for finding the absolute extremum of a function. Computational Mathematics and Mathematical Physics \textbf{12}, 57--67 (1972).
 \bibitem{c23} Shubert, B.O.: A sequential method seeking the global maximum of a function. SIAM Journal on Numerical Analysis \textbf{9}, 379--388 (1972).
-\bibitem{c24} Strongin R.G.: Multiextremal Minimization. Autom. Remote Control \textbf{7}, 1085--1088 (1970).
+\bibitem{c24} Strongin, R.G.: Multiextremal minimization. Autom. Remote Control \textbf{7}, 1085--1088 (1970).
 \bibitem{c25} Galperin, E.A.: The cubic algorithm. Journal of Mathematical Analysis and Applications \textbf{112}, 635--640 (1985).
@@
 \bibitem{c29} Sergeyev, Y.D.: Global one-dimensional optimization using smooth auxiliary functions. Mathematical Programming \textbf{81}, 127--146 (1998).
-\bibitem{c30} Sergeyev Ya.D., Grishagin V.A.: Sequential and parallel global optimization algorithms. Optimization Methods and Software, \textbf{3}, 111--124 (1994).
-\bibitem{c31} Sergeyev, Y.D.: An information global optimization algorithm with local tuning. SIAM J. Optim., \textbf{5}(4), 858--870 (1995).
-\bibitem{c32} Sergeyev, Y.D., Grishagin, V.A.: Parallel asynchronous global search and the nested optimization scheme. J. Comput. Anal. Appl., \textbf{3}(2), 123--145 (2001).
-\bibitem{c33} Lera D., Sergeyev Ya.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and Holder constants.
-Communications in Nonlinear Science and Numerical Simulation, \textbf{23}, 328--342 (2015).
-\bibitem{c34} Grishagin, V., Israfilov, R., Sergeyev, Y.: Comparative efficiency of dimensionality reduction schemes in global optimization. AIP Conference Proceedings 1776 (2016).
-\bibitem{c38} Gergel, V., Barkalov, K., Sysoyev, A.
-Globalizer: A novel supercomputer software system for solving time-consuming global optimization problem.
-Numerical algebra, control and optimization, \textbf{8}(1), 47--62 (2018).
+\bibitem{c30} Sergeyev, Ya.D., Grishagin, V.A.: Sequential and parallel global optimization algorithms. Optimization Methods and Software \textbf{3}, 111--124 (1994).
+\bibitem{c31} Sergeyev, Y.D.: An information global optimization algorithm with local tuning. SIAM J. Optim. \textbf{5}(4), 858--870 (1995).
+\bibitem{c32} Sergeyev, Y.D., Grishagin, V.A.: Parallel asynchronous global search and the nested optimization scheme. J. Comput. Anal. Appl. \textbf{3}(2), 123--145 (2001).
+\bibitem{c33} Lera, D., Sergeyev, Ya.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and Holder constants.
+Communications in Nonlinear Science and Numerical Simulation \textbf{23}, 328--342 (2015).
+\bibitem{c34} Grishagin, V., Israfilov, R., Sergeyev, Y.: Comparative efficiency of dimensionality reduction schemes in global optimization. AIP Conference Proceedings \textbf{1776}, 060011 (2016).
+\bibitem{c38} Gergel, V., Barkalov, K., Sysoyev, A.: Globalizer: a novel supercomputer software system for solving time-consuming global optimization problem.
+Numerical algebra, control and optimization \textbf{8}(1), 47--62 (2018).
 \bibitem{c39} {\v Z}ilinskas, A., Zilinskas, J.: Adaptation of a one-step worst-case optimal univariate algorithm of bi-objective Lipschitz optimization to multidimensional problems.
-Commun Nonlinear Sci Numer Simulat, \textbf{21}(1-3), 89--98 (2015).
+Communications in Nonlinear Science and Numerical Simulation \textbf{21}(1-3), 89--98 (2015).
 \doi{10.1016/j.cnsns.2014.08.025}
-\bibitem{c40} Evtushenko.
-Y., Posypkin, M.: A deterministic algorithm for global multiobjective optimization, Optimization Methods \& Software, \textbf{29}(5), 1005--1019 (2014).
+\bibitem{c40} Evtushenko, Y., Posypkin, M.: A deterministic algorithm for global multiobjective optimization.
+Optimization Methods \& Software \textbf{29}(5), 1005--1019 (2014).
 \doi{10.1080/10556788.2013.854357}
@@
 Remote Control \textbf{75}(6), 1025--1040 (2014).
-\bibitem{c42} Bleuler, S., Laumanns, M., Thiele, L., Zitzler, E.: Pisa-a platform and programming language independent interface for search algorithms, Evolutionary Multi-Criterion Optimization.
-LNCS, \textbf{2632}, 494--508 (2003).
+\bibitem{c42} Bleuler, S., Laumanns, M., Thiele, L., Zitzler, E.: PISA -- a platform and programming language independent interface for search algorithms.
+Evolutionary Multi-Criterion Optimization.
+LNCS \textbf{2632}, 494--508 (2003).
 \doi{10.1007/3-540-36970-8\_35}
-\bibitem{c43} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
-ACM Transactions on Mathematical Software, \textbf{29}(4), 469--480 (2003).
+\bibitem{c43} Gaviano, M., Kvasov, D.E, Lera, D., Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
+ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003).
 

### 2021-04-10 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
MOTOR правкти

@@
 In more complex situations, the objective functions that determine the effectiveness of decisions can be multiextremal, and decision making will require solving global optimization problems.
-In the most general case, however, the effectiveness of decisions may be determined by several objective criteria, which necessitates solving multi-objective optimization (MOO) problems.
+In the most general case, however, the effectiveness of decisions may be determined by several objective criteria, which makes it necessary to solve multi-objective optimization (MOO) problems.
 Thus, MOO problems are the most common decision making statements that one has to solve in many scientific and technical applications.
@@
 It is also important to note that by solving simultaneously a large number of scalar optimization problems thus generated it is possible to significantly decrease computational complexity of each separate problem.
-This effect is achived due to the fact that all such scalar problems  are based on the same MOO problem and, consequently, all computed values of the efficiency criteria of the MOO problem can be reduced to values of any scalar problem being solved simultaneously without any time consuming calculations.
+This effect is achieved due to the fact that all such scalar problems  are based on the same MOO problem and, consequently, all computed values of the efficiency criteria of the MOO problem can be reduced to values of any scalar problem being solved simultaneously without any time consuming calculations.
 In such cases, all the search information obtained in solving any single scalar problem can be used for solving all other scalar problems of the same set.
 
-This paper presents the results of the authors continuing research aimed to develop efficient methods for solving MOO problems \cite{c35,c36,c37,c44}.
+This paper presents the results of the authors' continuing research aimed to develop efficient methods for solving MOO problems \cite{c35,c36,c37,c44}.
 In \cite{c35}, some methods for solving global optimization problems are considered, in \cite{c36},  a general decision-making model based on MOO problems is proposed, and in \cite{c37}, a theoretical analysis is presented regarding the effectiveness of the use of search information when solving MOO problems.
@@
 
-The reference decision $y^0 \in D$ in (\ref{eq:05}) can be known \textit{a priori} or determined on a basis of some known prototype.
+The reference decision $y^0 \in D$ in (\ref{eq:05}) can be known \textit{a priori} or determined on the basis of some known prototype.
 In many cases, an abstract ideal decision $y^0 \in D$ is used as the reference decision $y^*\in D$ in which the objective functions $f_i$, $1 \leq i \leq s$, have minimal possible values, i.e.
@@
 At the same time, it should be noted that most works where dimensionality reduction is a key feature for solving multiextremal problems like (\ref{eq:04}) rely on the information-statistical theory of global search \cite{c15}.
-This theory has provided the a basis for developing a large number of efficient methods for multiextremal optimization \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
+This theory has provided the basis for developing a large number of efficient methods for multiextremal optimization \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.
 
@@
 
-Convergence conditions of the algorithms formulated in the form (\ref{eq:16})--(\ref{eq:18}) depend on the properties of the interval characteristics used.
+Convergence conditions of the algorithms are formulated in the form (\ref{eq:16})--(\ref{eq:18}) depend on the properties of the interval characteristics used.
 One of sufficient conditions for convergence of algorithms is, for example, the requirement that the characteristic of the interval containing the global minimum point should take on the maximum value at step 1 of the scheme (\ref{eq:16})--(\ref{eq:18}) during the global search iterations.
@@
 As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different efficient decisions due to possible changes in the optimality requirements.
-Obtaining different efficient decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multiextremal functions $F(\lambda, y)$ from (\ref{eq:04}).
+Obtaining different efficient decisions in the proposed approach is achieved by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multiextremal functions $F(\lambda, y)$ from (\ref{eq:04}).
 These functions $F(\lambda, y)$ can be optimized sequentially.
@@
 obtained in the calculations (see (\ref{eq:14})--(\ref{eq:17})).
-In (\ref{eq:22}) $x_i$, $1 \leq i \leq k$,  are the reduced points of performed global search iterations ordered in ascending order of coordinates,  $z_i$, $f_i$, $1 \leq i \leq k$, are values of the scalar function $F(\lambda, y)$, from (\ref{eq:04}) and objective functions $f_i(y)$, $1 \leq i \leq s$, from (\ref{eq:01}) of the current optimization problem to be solved at the points $x_i=y(x_i)$, $1 \leq i \leq k$.
+In (\ref{eq:22}) $x_i$, $1 \leq i \leq k$,  are the reduced points of performed global search iterations ordered in ascending order of coordinates,  $z_i$, $f_i$, $1 \leq i \leq k$, are the values of the scalar function $F(\lambda, y)$ from (\ref{eq:04}) and the objective functions $f_i(y)$, $1 \leq i \leq s$, from (\ref{eq:01}) of the current optimization problem to be solved at the points $x_i=y(x_i)$, $1 \leq i \leq k$.
 By using search information $A_k$ from (\ref{eq:22}) when choosing next search iterations it is possible to solve global optimization problems more efficiently and to provide convergence of algorithms only to the global minima of multiextremal functions being minimized.
@@
 
-For MGSA 100 problems (\ref{eq:04}) with different values of convolution coefficients $\lambda$ uniformly distributed in $\Lambda$ were solved.
+For MGSA, 100 problems (\ref{eq:04}) with different values of convolution coefficients $\lambda$ uniformly distributed in $\Lambda$ were solved.
 The results of the experiments performed are presented in Table \ref{tab:1}.
@@
 The columns named HV and DU contain the values of the HV and DU indicators.
-The results of the numerical experiments are divided with respect to the three algorithms used: GSA is described in Section \ref{sec:03}, MGSA is discussed in Section \ref{subsec:32}, SGSA is presented in Section \ref{subsec:43}.
+The results of the numerical experiments are divided with respect to the three algorithms used: GSA is described in Section \ref{sec:03}, MGSA is discussed in Subsection \ref{subsec:32}, SGSA is presented in Subsection \ref{subsec:43}.
 
@@
 
-The experimental results show that by reusing search information when optimizing 25 functions from the set $\Phi_p (y)$ the total computational iterations can be reduced  by more than 9.5 times without resorting to any additional computational resources.
+The experimental results show that by reusing search information when optimizing 25 functions from the set $\Phi_p (y)$, the total computational iterations can be reduced  by more than 9.5 times without resorting to any additional computational resources.
 At the same time, the MGSA algorithm performs the least number of iterations (the number of calculations of objective function values), while the SGSA algorithm provides a better approximation of the Pareto domain both in terms of approximation completeness and coverage uniformity.
@@
 When solving the problems, the accuracy $\varepsilon=0.05$ and the reliability $r=5.6$ were used.
-The results of numerical experiments are presented in Table \ref{tab:4} and Fig.\ref{fig:2}.
+The results of numerical experiments are presented in Table \ref{tab:4} and Fig.
+\ref{fig:2}.
 

### 2021-04-10 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
fix some mistakes in the reference list

@@
 
-\bibitem{c1} Miettinen, K.: Nonlinear Multiobjective Optimization. Springer (1999).
-\bibitem{c2} Ehrgott, M.: Multicriteria Optimization. Springer (2nd ed., 2010).
-\bibitem{c3} Collette, Y., Siarry, P.:  Multiobjective Optimization: Principles and Case Studies (Decision Engineering). Springer (2011).
-\bibitem{c4} Marler, R.T., Arora, J.S.: Multi-Objective Optimization: Concepts and Methods for Engineering. VDM Verlag (2009).
-\bibitem{c5} Pardalos, P.M., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-Convex Multi-Objective Optimization. Springer (2017).
-\bibitem{c6} Marler, R.T., Arora, J.S.: Survey of multi-objective optimization methods for engineering. Struct. Multidisciplinary Optimization \textbf{26}, 369--395 (2004).
-\bibitem{c7}  Figueira, J., Greco, S., Ehrgott, M., eds.: Multiple criteria decision analysis: state of the art surveys. New York (NY). Springer (2005).
-\bibitem{c8} Zavadskas, E.K., Turskis, Z., Kildien{\. e}, S.: State of art surveys of overviews on MCDM/MADM methods. Technological and Economic Development of Economy, \textbf{20}, 165--179 (2014).
-\bibitem{c9} Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications. Surv. Math. Ind. \textbf{11}, 1--42 (2005).
-\bibitem{c10} Collette, Y., Siarry, P.: Multiobjective Optimization: Principles and Case Studies. Decision Engineering. Springer, Heidelberg (2011). \doi{10.1007/978-3-662-08883-8}
-\bibitem{c11} Eichfelder, G.: Scalarizations for adaptively solving multi-objective optimization problems. Comput. Optim. Appl. \textbf{44}, 249--273 (2009).
+\bibitem{c1} Miettinen, K.: Nonlinear Multiobjective Optimization. Springer, Boston (1999)
+\bibitem{c2} Ehrgott, M.: Multicriteria Optimization, Springer, Berlin (2nd ed., 2010)
+\bibitem{c3} Collette, Y., Siarry, P.:  Multiobjective Optimization: Principles and Case Studies (Decision Engineering), Springer, Berlin  (2004)
+\bibitem{c4} Marler, R.T., Arora, J.S.: Multi-Objective Optimization: Concepts and Methods for Engineering, VDM Verlag (2009)
+\bibitem{c5} Pardalos, P.M., {\v Z}ilinskas, A., {\v Z}ilinskas, J.: Non-Convex Multi-Objective Optimization, Springer (2017)
+\bibitem{c6} Marler, R.T., Arora, J.S.: Survey of multi-objective optimization methods for engineering. Struct. Multidiscip. Optim. \textbf{26}, 369--395 (2004)
+\bibitem{c7}  Figueira, J., Greco, S., Ehrgott, M., eds.: Multiple Criteria Decision Analysis: State of the Art Surveys. Springer, New York (2005)
+\bibitem{c8} Zavadskas, E.K., Turskis, Z., Kildien{\. e}, S.: State of art surveys of overviews on MCDM/MADM methods. Technological and Economic Development of Economy, \textbf{20}, 165--179 (2014)
+\bibitem{c9} Hillermeier, C., Jahn, J.: Multiobjective optimization: survey of methods and industrial applications. Surv. Math. Ind. \textbf{11}, 1--42 (2005)
+\bibitem{c10} Collette, Y., Siarry, P.: Multiobjective Optimization: Principles and Case Studies. Decision Engineering. Springer, Heidelberg (2011) \doi{10.1007/978-3-662-08883-8}
+\bibitem{c11} Eichfelder, G.: Scalarizations for adaptively solving multi-objective optimization problems. Comput. Optim. Appl. \textbf{44}, 249--273 (2009)
 \bibitem{c12} Figueira, J., Liefooghe, A., Talbi, E., Wierzbicki, A.: A parallel multiple reference point approach for multi-objective optimization.
-European Journal of Operational Research \textbf{205}(2), 390--400 (2010).
-\doi{10.1016/j.ejor.2009.12.027}
+European Journal of Operational Research \textbf{205}(2), 390--400 (2010) \doi{10.1016/j.ejor.2009.12.027}
 \bibitem{c35} Gergel, V.: An unified approach to use of coprocessors of various types for solving global optimization problems.
-2nd International Conference on Mathematics and Computers in Sciences and in Industry, 13--18 (2015).
-\doi{10.1109/MCSI.2015.18}
+2nd International Conference on Mathematics and Computers in Sciences and in Industry, 13--18 (2015) \doi{10.1109/MCSI.2015.18}
 \bibitem{c36} Gergel, V.P., Kozinov, E.A.: Accelerating multicriterial optimization by the intensive exploitation of accumulated search data.
-AIP Conference Proceedings \textbf{1776}, 090003 (2016).
-\doi{10.1063/1.4965367}
-\bibitem{c37} Gergel, V., Kozinov, E.: Efficient multicriterial optimization based on intensive reuse of search information. Journal of Global Optimization \textbf{71}(1), 73--90 (2018).
+AIP Conference Proceedings \textbf{1776}, 090003 (2016) \doi{10.1063/1.4965367}
+\bibitem{c37} Gergel, V., Kozinov, E.: Efficient multicriterial optimization based on intensive reuse of search information. J. Glob. Optim. \textbf{71}(1), 73--90 (2018)
 \bibitem{c44} Gergel, V., Kozinov, E., Barkalov, K.: Computationally efficient approach for solving lexicographic multicriteria optimization problems.
@@
 Lett.
-(2020).
-\doi{10.1007/s11590-020-01668-y}
-\bibitem{c13} Zhigljavsky, A.A.: Theory of Global Random Search. Dordrecht: Kluwer Academic Publishers (1991).
-\bibitem{c14} Pint{\' e}r, J.D.: Global optimization in Action (continuous and Lipschitz optimization: algorithms, implementations and applications). Kluwer Academic Publishers, Dortrecht (1996).
-\bibitem{c15} Strongin, R., Sergeyev, Ya.: Global optimization with non-convex constraints. Sequential and parallel algorithms. Kluwer Academic Publishers, Dordrecht (2nd ed. 2013, 3rd ed. 2014).
-\bibitem{c16} Yang, X.-S.: Nature-inspired metaheuristic algorithms. Luniver Press, Frome (2008).
-\bibitem{c17} Locatelli, M., Schoen, F.: Global Optimization: Theory, Algorithms, and Applications. SIAM (2013).
+(2020) \doi{10.1007/s11590-020-01668-y}
+\bibitem{c13} Zhigljavsky, A.A.: Theory of Global Random Search.  Kluwer Academic Publishers, Dordrecht (1991).
+\bibitem{c14} Pint{\' e}r, J.D.: Global Optimization in Action (Continuous and Lipschitz Optimization: Algorithms, Implementations and Applications). Kluwer Academic Publishers, Dortrecht (1996)
+\bibitem{c15} Strongin, R., Sergeyev, Ya.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2nd ed. 2013, 3rd ed. 2014)
+\bibitem{c16} Yang, X.-S.: Nature-inspired Metaheuristic Algorithms. Luniver Press, Frome (2008)
+\bibitem{c17} Locatelli, M., Schoen, F.: Global Optimization: Theory, Algorithms, and Applications. SIAM (2013)
 \bibitem{c18} Paulavi{\v c}ius, R., {\v Z}ilinskas, J.: Simplicial Global Optimization. Springer, New York (2014).
-\bibitem{c19} Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. Princeton University Press (2016).
-\bibitem{c20} Sergeyev, Y.D., Kvasov, D.E.: Deterministic Global Optimization: An Introduction to the Diagonal Approach. Springer, New York (2017). 
-\bibitem{c21} Sergeyev, Y.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013).
-\bibitem{c22} Piyavskij, S.: An algorithm for finding the absolute extremum of a function. Computational Mathematics and Mathematical Physics \textbf{12}, 57--67 (1972).
-\bibitem{c23} Shubert, B.O.: A sequential method seeking the global maximum of a function. SIAM Journal on Numerical Analysis \textbf{9}, 379--388 (1972).
-\bibitem{c24} Strongin, R.G.: Multiextremal minimization. Autom. Remote Control \textbf{7}, 1085--1088 (1970).
-\bibitem{c25} Galperin, E.A.: The cubic algorithm. Journal of Mathematical Analysis and Applications \textbf{112}, 635--640 (1985).
-\bibitem{c26} Breiman, L., Cutler, A.: A deterministic algorithm for global optimization, Mathematical Programming \textbf{58}, 179--199 (1993).
-\bibitem{c27} Baritompa, W.: Accelerations for a variety of global optimization methods. Journal of Global Optimization \textbf{4}, 37--45 (1994).
-\bibitem{c28} Gergel, V.P.: A method of using derivatives in the minimization of multiextremum functions. Computational Mathematics and Mathematical Physics \textbf{36}(6), 729--742 (1996).
-\bibitem{c29} Sergeyev, Y.D.: Global one-dimensional optimization using smooth auxiliary functions. Mathematical Programming \textbf{81}, 127--146 (1998).
-\bibitem{c30} Sergeyev, Ya.D., Grishagin, V.A.: Sequential and parallel global optimization algorithms. Optimization Methods and Software \textbf{3}, 111--124 (1994).
-\bibitem{c31} Sergeyev, Y.D.: An information global optimization algorithm with local tuning. SIAM J. Optim. \textbf{5}(4), 858--870 (1995).
-\bibitem{c32} Sergeyev, Y.D., Grishagin, V.A.: Parallel asynchronous global search and the nested optimization scheme. J. Comput. Anal. Appl. \textbf{3}(2), 123--145 (2001).
+\bibitem{c19} Floudas, C.A., Pardalos, M.P.: Recent Advances in Global Optimization. Princeton University Press (2016)
+\bibitem{c20} Sergeyev, Y.D., Kvasov, D.E.: Deterministic Global Optimization: An Introduction to the Diagonal Approach. Springer, New York (2017) 
+\bibitem{c21} Sergeyev, Y.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting Space-Filling Curves. Springer (2013)
+\bibitem{c22} Piyavskij, S.: An algorithm for finding the absolute extremum of a function. Comput. Math. Math. Phys. \textbf{12}, 57--67 (1972)
+\bibitem{c23} Shubert, B.O.: A sequential method seeking the global maximum of a function. SIAM J. Numer. Anal. \textbf{9}, 379--388 (1972)
+\bibitem{c24} Strongin, R.G.: Multiextremal minimization. Autom. Remote Control \textbf{7}, 1085--1088 (1970)
+\bibitem{c25} Galperin, E.A.: The cubic algorithm. J. Math. Anal. Appl. \textbf{112}, 635--640 (1985)
+\bibitem{c26} Breiman, L., Cutler, A.: A deterministic algorithm for global optimization. Math. Program. \textbf{58}, 179--199 (1993)
+\bibitem{c27} Baritompa, W.: Accelerations for a variety of global optimization methods. J. Glob. Optim. \textbf{4}, 37--45 (1994)
+\bibitem{c28} Gergel, V.P.: A method of using derivatives in the minimization of multiextremum functions. Comput. Math. Math. Phys. \textbf{36}(6), 729--742 (1996)
+\bibitem{c29} Sergeyev, Y.D.: Global one-dimensional optimization using smooth auxiliary functions. Math. Program. \textbf{81}, 127--146 (1998)
+\bibitem{c30} Sergeyev, Ya.D., Grishagin, V.A.: Sequential and parallel global optimization algorithms. Optim. Methods Softw.\textbf{3}, 111--124 (1994)
+\bibitem{c31} Sergeyev, Y.D.: An information global optimization algorithm with local tuning. SIAM J. Optim. \textbf{5}(4), 858--870 (1995)
+\bibitem{c32} Sergeyev, Y.D., Grishagin, V.A.: Parallel asynchronous global search and the nested optimization scheme. J. Comput. Anal. Appl. \textbf{3}(2), 123--145 (2001)
 \bibitem{c33} Lera, D., Sergeyev, Ya.D.: Deterministic global optimization using space-filling curves and multiple estimates of Lipschitz and Holder constants.
-Communications in Nonlinear Science and Numerical Simulation \textbf{23}, 328--342 (2015).
-\bibitem{c34} Grishagin, V., Israfilov, R., Sergeyev, Y.: Comparative efficiency of dimensionality reduction schemes in global optimization. AIP Conference Proceedings \textbf{1776}, 060011 (2016).
+Communications in Nonlinear Science and Numerical Simulation \textbf{23}, 328--342 (2015)
+\bibitem{c34} Grishagin, V., Israfilov, R., Sergeyev, Y.: Comparative efficiency of dimensionality reduction schemes in global optimization. AIP Conference Proceedings \textbf{1776}, 060011 (2016)
 \bibitem{c38} Gergel, V., Barkalov, K., Sysoyev, A.: Globalizer: a novel supercomputer software system for solving time-consuming global optimization problem.
-Numerical algebra, control and optimization \textbf{8}(1), 47--62 (2018).
+Numerical Algebra, Control and Optimization \textbf{8}(1), 47--62 (2018)
 \bibitem{c39} {\v Z}ilinskas, A., Zilinskas, J.: Adaptation of a one-step worst-case optimal univariate algorithm of bi-objective Lipschitz optimization to multidimensional problems.
-Communications in Nonlinear Science and Numerical Simulation \textbf{21}(1-3), 89--98 (2015).
-\doi{10.1016/j.cnsns.2014.08.025}
-\bibitem{c40} Evtushenko, Y., Posypkin, M.: A deterministic algorithm for global multiobjective optimization.
-Optimization Methods \& Software \textbf{29}(5), 1005--1019 (2014).
-\doi{10.1080/10556788.2013.854357}
+Communications in Nonlinear Science and Numerical Simulation \textbf{21}(1-3), 89--98 (2015) \doi{10.1016/j.cnsns.2014.08.025}
+\bibitem{c40} Evtushenko, Y., Posypkin, M.: A deterministic algorithm for global multiobjective optimization. Optim. Methods Softw. \textbf{29}(5), 1005--1019 (2014) \doi{10.1080/10556788.2013.854357}
 \bibitem{c41} Evtushenko, Y., Posypkin, M.: Method of non-uniform coverages to solve the multicriteria optimization problems with guaranteed accuracy.
 Autom.
-Remote Control \textbf{75}(6), 1025--1040 (2014).
+Remote Control \textbf{75}(6), 1025--1040 (2014)
 \bibitem{c42} Bleuler, S., Laumanns, M., Thiele, L., Zitzler, E.: PISA -- a platform and programming language independent interface for search algorithms.
-Evolutionary Multi-Criterion Optimization.
-LNCS \textbf{2632}, 494--508 (2003).
-\doi{10.1007/3-540-36970-8\_35}
+Lecture Notes in Computer Science \textbf{2632}, 494--508 (2003) \doi{10.1007/3-540-36970-8\_35}
 \bibitem{c43} Gaviano, M., Kvasov, D.E, Lera, D., Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization.
-ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003).
+ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003)
 

### 2021-04-12 2021/RSD_Ufa/RSD2021.tex
Our part for the RSD paper

%%%%%%%%%%%%%%%%%%%% author.tex 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{svproc}
%
% RECOMMENDED 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%
\usepackage{graphicx}
\usepackage{marvosym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cite}

%\usepackage[russian]{babel}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\usepackage{hyperref}
\def\UrlFont{\rmfamily}

\def\orcidID#1{\unskip$^{[#1]}$}
\def\letter{$^{\textrm{(\Letter)}}$}

\begin{document}
\mainmatter              % start of a contribution
%
\title{Parallel Global Search Algorithm for Optimization of the Kinetic Parameters of Chemical Reactions}
%
\titlerunning{Adaptive Global Optimization}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{
Irek Gubaidullin$^{1}$\and
Leniza Enikeeva$^{1}$\and
Konstantin Barkalov$^2$%\orcidID{0000-0001-5273-2471}
\and
Ilya Lebedev$^2$%\orcidID{0000-0002-8736-0652} 
}

%
\authorrunning{I. Gubaidullin et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Konstantin Barkalov and Ilya Lebedev }
%

\institute{$^2$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
\email{konstantin.barkalov@itmm.unn.ru},
\email{ilya.lebedev@itmm.unn.ru}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
The paper considers the application of parallel computing technology to the simulation of a catalytic chemical reaction, which is widely used in the modern chemical industry to produce synthesis gas. As a chemical reaction, the process of pre-forming propane on a Ni catalyst is assumed. To simulate a chemical process, it is necessary to develop a kinetic model of the process, that is, to determine the kinetic parameters. To do this, the inverse problem of chemical kinetics is solved, which predicts the values of kinetic parameters based on laboratory data. From a mathematical point of view, the inverse of the chemical kinetics problem is a global optimization problem. A parallel information-statistical global search algorithm was used to solve it. The use of the parallel algorithm has reduced significantly the search time for the optimum. The found optimal parameters of the model made it possible to adequately simulate the process of pre-forming propane on a Ni-catalyst.

\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing }
\end{abstract}

\section{Introduction}

\section{Parallel Global Search Algorithm}\label{Sec_GSA}

Let us consider global optimization problem of the form 
\begin{gather}
 \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problem}\\
 D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \; a_i,b_i\in R, \;  1\leq i \leq N\right\} \label{D},
\end{gather}
where the objective function is a black-box function and it is assumed to satisfy the Lipschitz condition
\[
\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D,
\]
with the constant $L, \; L<\infty,$ unknown a priori.

The assumption of the objective function to be Lipschitzian is typical of many approaches to the development of the deterministic global optimization algorithms.
The first methods of Lipschitz optimization were proposed in the early 1970s \cite{Piyavskii1972,Shubert1972}. Since that time, this line of research has continued to develop actively \cite{Evtushenko2013,Zilinskas2010,Pinter1996,Jones2009}.

Currently, nature-inspired algorithms are widely used for solving optimization problems with the black-box objective functions; for example, see \cite{Yang2013,Gendreau2010,Eiben2015}. Such algorithms, in one way or another, employ the ideas of random search. Due to simple implementation and use, they have become very popular. However, these algorithms are inferior to the deterministic counterparts in terms of quality \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a certain
set).

One of the efficient deterministic methods for solving multiextremal optimization problems is \textit{the information-statistical global search algorithm} \cite{Strongin2000}. This method initially proposed for solving unconstrained optimization problems was successfully generalized to the classes of optimization problems with non-convex constraints and multicriteria optimization problems. For different versions of the algorithm, parallelization methods taking into account the architecture of modern computing systems were also suggested \cite{Barkalov2016,globalizerSystem,Strongin2018}.%поставить ссылку

The parallel global search algorithm described in this section is based on the idea of reducing the dimension with the use of the Peano curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously map the unit interval $[0,1]$ onto the $N$-dimensional cube $D$ from (\ref{D}). By using this kind of mapping, it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem
\[
\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}},
\]
where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
\[
\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}
\]
with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation
$ H=2 L \sqrt{N+3}$ and $y(x)$ is a Peano curve from $[0,1]$ onto $D$.

Note that theoretically the Peano curve $y(x)$ is defined as a limit object. Therefore, in practical implementation, only some approximation to the true space-filling curve can be constructed. Some methods for constructing this type of approximations (called \textit{evolvents}) are considered in \cite{Sergeyev2013,Strongin2000}. In this case, the accuracy of the approximation to the true curve $y(x)$ depends on the density of the evolvent $m$ (which is a parameter for constructing the evolvent) and is of the order of $2^{-m}$ for each coordinate.
Examples of evolvents with different values of $m$  for the dimensionality $N=3$ are shown in Fig.~\ref{evolvents}.

\begin{figure}[ht]
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1a.JPG} \\ (a)}
\end{minipage}
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (b)}
\end{minipage}
\caption{Evolvents in three dimensions with (a) $M=3$ and (b) $M=4$}
\label{evolvents}
\end{figure}

Let us call the process of computing a function value (including the construction of the image $y=y(x)$) as a \textit{trial}, and the pair $\{x, z = \varphi(y(x))\}$ as the outcome of the trial.
When describing a parallel algorithm, we will use the term \textit {iteration} to denote the simultaneous (parallel) execution of several trials (each trial is executed by a separate processor). The number of trials during the $n$-th iteration we will denote as $p$ , and the total number of trials executed during all $n$ iterations as $k(n)$ . In other words, we assume that we have  $p$  processors at our disposal while executing the $n$-th iteration.

The parallel global search algorithm used in this research (according to \cite{Strongin2000}) can be formulated as follows.
The first $p$ trials are executed at the points $x^0 = 0$, $x^1 = 1$ and at the arbitrary internal points $x^2, ..., x^{p-1}$ of the interval $(0,1)$. Let us assume $n \geq 1$  iterations of the method to be completed, in the course of which the trials in $k=k(n)$ points $x^i, 0 \leq i \leq k,$ have been executed. Then, the points $x^{k+1},...,x^{k+p}$  of the search trials for the next $(n+1)$-th iteration are determined according to the following rules.

\begin{enumerate}
	\item 
	Renumber the inverse images of all the points from the trials already performed  
\begin{equation}\label{y_i}
y^0=y(x^0), y^1=y(x^1),...,y^k=y(x^k)
\end{equation}
by subscripts in the increasing order of their coordinates, i.e.
\begin{equation}\label{x_i}
0=x_0<x_1<\dots <x_k=1,
\end{equation}
and associate these with the values $z_i=\varphi(y(x_i)), 0\leq i \leq k,$  computed at these points.
\item
Compute the maximum absolute value of the first divided differences
\begin{equation}\label{mu}
\mu = \max_{1 \leq i \leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
\end{equation}
where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$. If $\mu = 0$, set $\mu = 1$.
\item
For each interval $(x_{i-1}, x_i), \; 1\leq i \leq k,$  calculate the value
\begin{equation}\label{R}
R(i)=r\mu\Delta_i+\frac{(z_i-z_{i-1})^2}{r\mu\Delta_i}-2(z_i+z_{i-1})
\end{equation}
called the \textit{characteristic} of the interval; the real number $r>1$ being the input parameter of the algorithm.

\item 
Arrange characteristics  $R(i), 1 \leq i \leq k$, in decreasing order 
\begin{equation}\label{Rdec}
R(t_1)\geq R(t_2)\geq \dots \geq R(t_{k})
\end{equation}
and select $p$ largest characteristics with interval numbers $t_j, 1\leq j \leq p$.
\item
Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), 1\leq j\leq p$, computed according to the formula
\[
x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - \mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N.
\]

\end{enumerate}

The algorithm stops if condition $\Delta_{t_j}<\epsilon$ is satisfied for at least one index $t_j, 1 \leq j \leq p$ ; here $\epsilon>0$ is the predefined accuracy. 
%Также алгоритм останавливается, если превышено заданное при запуске число испытаний $K_{max}$.

The considered variant of the parallelization is \textit{synchronous}, when the transition to the next iteration is performed after full termination of the current one, i.e., after completing the last trial of the current iteration.
% Конкретная реализация данной схемы распараллеливания будет зависеть от архитектуры вычислительной системы, а также от требований к программному обеспечению, необходимому для вычисления значений целевой функции задачи. В данном исследовании проведение испытаний требовало решения жесткой системы ODE (для ее решения был использован метод Radau IIA из бибилиотеки SciPy языка Python). Поэтому распараллеливание было ораганизовано на CPU с использованием технологии MPI. В одном (master) процессе работал алгоритм глобального поиска; остальные (slave) процессы проводили параллельные испытания в рамках выполнения шага 5 алгорита. Объем передаваемых данных между процессами будет небольшим: требуется лишь разослать координаты точек испытаний и затем собрать вычисленные значения целевой функции в этих точках. Одновременно с этим процесс проведения испытания достаточно трудоемкий (время проведения одного испытания занимает не менее 1 сек.), что значительно превышает время передачи данных между процессами.

\section{Numerical Experiments}\label{Sec_Exp}

The numerical experiments were carried out using the Lobachevsky supercomputer (Lobachevsky State University of Nizhny Novgorod). 
%Указать актуальную конфигурацию
%Each supercomputer node included two quad-core processors Intel Xeon X5570, two NVIDIA Tesla X2070 and 12 Gb RAM.
%Указать актуальное программное обеспечение
%To build the program for running on the Lobachevsky supercomputer, the GCC 4.3.0 compiler, CUDA 6.5 and Intel MPI 2017 were used.
%Для вычисления значений целевой функции задачи использовался Python .

%Parallel global search algorithm запускался со следующими параметрами: параметр надежности $r=3$, точность поиска $\epsilon = 0.01$, плотность построения развертки $m=10$, максимальное число испытаний $K_{max}=1000$. После завершения работы parallel global search algorithm найденное решение уточнялось с помощью локального поиска, использовался Hooke--Jeeves method \cite{HookJeeves}. Точность локального поиска составляла $\epsilon = 0.001$.

%Далее - описать параллельный запуск и его результаты, например, в виде таблицы с ускорением

%Затем - проинтерпретировать результаты с содержателной точки зрения

\section{Conclusion}

%планы дальнейших работ - реализация метода решения системы

\medskip

\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204.

%
% ---- Bibliography ----
%
\bibliographystyle{spmpsci}
\bibliography{bibliography}{}

\end{document}

### 2021-04-13 2021/RSD_Ufa/RSD2021.tex
include two figures

@@
 
-%\usepackage[russian]{babel}
+\usepackage[russian]{babel}
 
@@
 \author{
-Irek Gubaidullin$^{1}$\and
-Leniza Enikeeva$^{1}$\and
-Konstantin Barkalov$^2$%\orcidID{0000-0001-5273-2471}
+Irek Gubaydullin$^{1,2}$\and
+Leniza Enikeeva$^{2,3}$\orcidID{0000-0003-4219-4870}
 \and
-Ilya Lebedev$^2$%\orcidID{0000-0002-8736-0652} 
+Konstantin Barkalov$^4$\orcidID{0000-0001-5273-2471}
+\and
+Ilya Lebedev$^4$\orcidID{0000-0002-8736-0652} 
 }
@@
 %
-\authorrunning{I. Gubaidullin et al.} % abbreviated author list (for running head)
+\authorrunning{I. Gubaydullin et al.} % abbreviated author list (for running head)
 %
@@
 
-\institute{$^2$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
+\institute{$^1$Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia\\$^2$Ufa State Petroleum Technological University, Ufa, Russia \\$^3$Novosibirsk State University, Novosibirsk, Russia\\$^4$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
+\email{leniza.enikeeva@yandex.ru},
 \email{konstantin.barkalov@itmm.unn.ru},
@@
 The paper considers the application of parallel computing technology to the simulation of a catalytic chemical reaction, which is widely used in the modern chemical industry to produce synthesis gas.
-As a chemical reaction, the process of pre-forming propane on a Ni catalyst is assumed.
+As a chemical reaction, the process of pre-reforming propane on a Ni catalyst is assumed.
 To simulate a chemical process, it is necessary to develop a kinetic model of the process, that is, to determine the kinetic parameters.
@@
 The use of the parallel algorithm has reduced significantly the search time for the optimum.
-The found optimal parameters of the model made it possible to adequately simulate the process of pre-forming propane on a Ni-catalyst.
+The found optimal parameters of the model made it possible to adequately simulate the process of pre-reforming propane on a Ni-catalyst.
 
@@
 \section{Introduction}
+
+%Одним из направлений лаборатории математической химии Института нефтехимии и катализа УФИЦ РАН является исследование химических реакций методами математического моделирования.
+На данный момент изучено большое количество сложных химических процессов -- реакция циклоалюминирования алкенов триэтилалюминием в алюмациклопентаны, реакция гидроалюминирования олефинов алюминийорганическими соединениями, реакция спиртов с диметилкарбонатом и др.
+Математическое описание химических реакций необходимо для построения математических моделей химической кинетики, которые называются кинетическими моделями.
+Кинетическая модель является основой математического моделирования химических реакций и исходным уровнем при математическом моделировании химических реакторов.
+При разработке кинетической модели многостадийной химической реакции решается обратная задача химической кинетики (ХК).
+Обратной задачей ХК в общем случае называют процесс выбора механизма сложного процесса и определение его количественных характеристик -- констант скорости одностадийных реакций на основе опытных данных о динамике протекания процесса.
+С математической точки зрения обратная задача химической кинетики представляет собой задачу оптимизации, поэтому актуальной является задача выбора метода оптимизации.
+
+One of the areas of the Laboratory of Mathematical Chemistry of the Institute of Petrochemistry and Catalysis of the UFIC RAS is the study of chemical reactions by mathematical modeling methods.
+At the moment, a large number of complex chemical processes have been studied-the reaction of cycloalumination of alkenes with triethylaluminium into alumacyclopentanes, the reaction of hydroalumination of olefins with organoaluminum compounds, the reaction of alcohols with dimethyl carbonate, etc.
+The mathematical description of chemical reactions is necessary for the construction of mathematical models of chemical kinetics, which are called kinetic models.
+The kinetic model is the basis for mathematical modeling of chemical reactions and the initial level for mathematical modeling of chemical reactors.
+When developing a kinetic model of a multi-stage chemical reaction, the inverse problem of chemical kinetics (CK) is solved.
+The inverse problem of HC is generally called the process of selecting the mechanism of a complex process and determining its quantitative characteristics - the rate constants of single-stage reactions based on experimental data on the dynamics of the process.
+From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem, so the problem of choosing an optimization method is relevant.
+
+%Russian
+Метод решения оптимизационных задач подобного типа должен учитывать следующую специфику.
+Во-первых, целевая функция задачи является "черным ящиком", т.к.
+она задается не в виде формулы, а в виде некоторого алгоритма вычисления ее значений.
+Во-вторых, вычисление даже одного значения целевой функции является трудоемкой операцией, т.к.
+требует численного моделирования.
+Перечисленная специфика не позволяет задействовать для решения обратных задач мультистартовые методы, основанные на идеях многократного градиентного спуска.
+С одной стороны, вычисление градиента в данном случае будет чрезмерно дорогой операцией, а с другой стороны - мультистартовые методы слабо применимы для существенно многоэкстремальных задач, какими и являются обратные задачи химической кинетики.
+
+Перечисленные особенности inverse problem of chemical kinetics учитываются в разработанном в University of Nizhni Novgorod параллельном алгоритме глобальной оптимизации.
+Теоретическим базисом данного алгоритма является information-statistical approach, детально изложенный в \cite{Strongin2000}.
+Разработанный алгоритм основан на редукции исходной многомерной задачи к эквивалентной ей одномерной задаче с последующим ее решением эффективными методами оптимизации функций одной переменной.
+Вопросы распараллеливания алгоритма на разных архитектурах рассмотрены в \cite{Barkalov2016}.
+В данной статье представлены результаты, полученные при использовании parallel global search algorithm для решения inverse problem of chemical kinetics.
+
+\section{Mathematical Model}\label{Sec_math_mod}
+
+The process under study is propane pre-reforming into methane-rich gas over Ni catalyst, which is an industrially important chemical process \cite{USKOV2019126741, Uskov_en}.
+Pre-reforming of propane was studied over industrial nickel-chromium catalyst at pressure of 1 and 5 bar in the temperature range of 220 -- 380 $^0$C and at flow rates of 4000 and 12000 h$^{-1}$.
+The experimental data on propane pre-reforming was acquired from our previous work \cite{Uskov_en}.
+The reaction scheme consists of two reactions: propane steam conversion and CO$_2$ methanation \cite{Uskov_cat}:
+\begin{gather}
+C_3H_8 + 6H_2O \rightarrow 10H_2 + 3CO_2 \label{reac_ref}\\
+CO_2 + 4H_2 \rightleftharpoons CH_4 + 2H_2O \label{reac_met},
+\end{gather}
+
+where the rates of reactions \eqref{reac_ref} -- \eqref{reac_met} are expressed according to the Langmuir-Hinshelwood model \cite{USKOV2019126741}:
+
+\[
+W_{ref} = \dfrac{k_{ref}^0 \cdot \exp \left(- \dfrac{E_{ref}}{RT} \right) }{\left(1 + B \cdot C_{C3H8}\right)^m},
+\]
+
+\[
+W_{met} = k_{met}^0 \cdot \exp \left(- \dfrac{E_{met}}{RT} \right) \cdot C_{H2} \cdot \left[1 - \dfrac{P_{CH4} P_{H2O}^2}{K_{eq} P_{CO2} P_{H2}^4 }\right] ,
+\]
+
+where $W_{ref}$ and $W_{met}$ are the reaction rates; $E_{ref}$ and $E_{met}$ are the observed activation energies, J/mol; $k^0_{ref}$ and $k^0_{met}$ are the pre–exponential multipliers; $B$ is the constant parameter, $T$ is the temperature, K; $R$ is the universal gas constant, J(K$\cdot$ mol).
+The "$ref$"and "$met$" indexes refer to pre-reforming and methanation reactions, respectively.
+$C_{C3H8}$ and $C_{H2}$ are concentrations of propane and hydrogen, mol/m$^3$; $m$ is order of the denominator, which varied from 0 to 2; $K_{eq}$ is the equilibrium constant of $CO_2$ methanation; $P_{CH4}$, $P_{H2O}$, $P_{CO2}$, $P_{H2}$ are partial pressures of the corresponding substances, bar.
+The mathematical model is a system of equations of material balance:
+
+\begin{equation}\label{mater_balance}
+\begin{cases}
+G \dfrac{dy_i}{dl} = \left(\nu_i^{ref} W_{ref} + \nu_i^{met} W_{met} \right) m_i, 
+\\
+0 \le l \le , i \in \left\lbrace C_3H_8, CH_4, H_2O, H_2, CO_2 \right\rbrace,
+\\
+l = 0, y_i = y_{i0},
+\end{cases}
+\end{equation}
+
+where $G$ is a mass flow of the mixture, kg/(m$^2 \cdot$ sec); $y_i$ is a mass fraction of the $i$-th component; $\nu_i$ is a stoichiometric coefficient of the $i$-th component; $m_i$ is a molar mass of the $i$-th component, kg/mol; $l$ is coordinate along the catalytic layer, m; $L$ is a length of the catalytic layer, m.
+The length of the catalytic layer is 0.008 m.
+The mathematical model of chemical kinetics problems is a system of differential equations that describes the variations in substance concentrations over time according to the rates of reaction stages.
+The system of differential equations is a Cauchy problem containing the initial data \cite{ENIKEEV2020123131, Chainikova}.
+The numerical solving of such a system of equations is a direct problem of chemical kinetics.
+
+%Системы обыкновенных дифференциальных уравнений (ОДУ), описывающие химические прцоессы, зачастуя являются жесткими за счет наличия быстропротекающих и медленнопротекающих стадий реакции.
+Поэтому в качестве метода решения системы ОДУ был выбран метод RADAU-IIA, который подходит для решения жестких систем ОДУ.
+Рассмотрим в общем случае задачу Коши вида:
+
+Systems of ordinary differential equations (ODES) describing chemical processes are often rigid due to the presence of fast-flowing and slow-flowing reaction stages.
+Therefore, the RADAU-IIA method was chosen as a method for solving the ODE system, which is suitable for solving rigid ODE systems.
+Let us consider in the general case the Cauchy problem of the form:
+
+\begin{equation}\label{ode}
+\dfrac{dx}{dt} = f(\textbf{x}, y), \textbf{x}(0) = \textbf{x}_0, 0 \le t \le T,
+\end{equation}
+
+%где x в нашем случае это вектор концентраций веществ реакции размерности 5, а f это правая часть уравнения \eqref{mater_balance}.
+Тогда s-стадийный метод Рунге-Кутты для решения системы \eqref{ode} выписывается следующим образом:
+
+where $x$ in our case is the vector of concentrations of reaction substances of dimension 5, and $f$ is the right-hand side of the equation \eqref{mater_balance}.
+Then the $s$-stage Runge-Kutta method for solving the \eqref{ode} system is written as follows:
+
+\begin{equation}\label{runge}
+\begin{cases}
+X_i = x_n + h \sum\limits_{j=1}^s a_{ij} f(X_j, t_n + c_j h) , 1 \le i \le s, 
+\\
+x_{n+1} = x_n + h \sum\limits_{i=1}^s b_i f (X_i, t_n + c_i h),
+\end{cases}
+\end{equation}
+
+%где $x_n$, $x_{n+1}$, $X_i$ -- приближенные решения системы \eqref{ode} в моменты времени $t_n$, $t_{n+1} = t_n + h$, $t_n + c_i h$.
+Коэффициенты  методов  Рунге-Кутты  обычно  записывают  в  виде таблицы Бутчера:
+
+where $x_n$, $x_{n+1}$, $X_i$ are approximate solutions of the \eqref{ode} system at time points $t_n$, $t_{n+1} = t_n + h$, $t_n + c_i h$.
+The coefficients of the Runge-Kutta methods are usually written in the form of a Butcher table:
+
+\begin{equation}\label{butcher}
+\begin{array}{c|cccc}
+c_1 & a_{11} & a_{12} & ... & a_{1s} \\
+c_2 & a_{21} & a_{22} & ... & a_{2s} \\
+... & ... & ... & ... & ... \\
+c_s & a_{s1} & a_{s2} & ... & a_{ss} \\
+\hline
+ & b_{1} & b_{2} & ... & b_{s} \\
+\end{array}
+\end{equation}
+
+% Таблица  Бутчера  содержит  векторы  b  и  с,  состоящие  из коэффициентов  bi  и  ci  соответственно  и  матрицу  A  размерностью $s \times s$, включающую  коэффициенты  aij. 
+%Метод Радо-IIA пятого порядка (Radau 5):
+
+The Butcher table contains the vectors $b$ and $c$, consisting of the coefficients $b_i$ and $c_i$, respectively, and the matrix $A$ of dimension $s \times s$, including the coefficients $a_{ij}$.
+
+Fifth-order Rado-IIA method (Radau 5):
+
+\begin{equation}\label{radau}
+\begin{array}{c|ccc}
+\dfrac{4 - \sqrt{6}}{10} & \dfrac{88 - 7\sqrt{6}}{360} & \dfrac{296 - 169\sqrt{6}}{1800} & \dfrac{-2 + 3\sqrt{6}}{225} \\
+\dfrac{4 + \sqrt{6}}{10} & \dfrac{296 + 169\sqrt{6}}{1800} & \dfrac{88 + 7\sqrt{6}}{360} & \dfrac{-2 - 3\sqrt{6}}{225} \\
+1 & \dfrac{16 - \sqrt{6}}{36} & \dfrac{16 + \sqrt{6}}{36} & \dfrac{1}{9} \\
+\hline
+& \dfrac{16 - \sqrt{6}}{36} & \dfrac{16 + \sqrt{6}}{36}  & \dfrac{1}{9} \\
+\end{array}
+\end{equation}
+
+Determining the kinetic parameters of reaction stages by comparing calculated values of substance concentrations and experimental results is an inverse problem of chemical kinetics.
+The mathematical problem is to minimize the functional of the deviation between calculated and experimental values.
+The functional of minimizations determined as the sum of absolute deviations between calculated and experimental concentrations:
+
+\begin{equation}\label{func}
+F = \sum\limits_{i=1}^M \sum\limits_{j=1}^N \left| x_{ij}^{calc} - x_{ij}^{exp} \right| \rightarrow \min,
+\end{equation}
+
+where $x_{ij}^{calc}$ and $x_{ij}^{exp}$ are calculated and experimental values of component concentrations; $M$ is the number of measuring points; $N$ is the number of substances involved in the reaction.
 
@@
 Due to simple implementation and use, they have become very popular.
-However, these algorithms are inferior to the deterministic counterparts in terms of quality \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a certain
-set).
+However, these algorithms are inferior to the deterministic counterparts in terms of quality \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a certain set).
 
@@
 The algorithm stops if condition $\Delta_{t_j}<\epsilon$ is satisfied for at least one index $t_j, 1 \leq j \leq p$ ; here $\epsilon>0$ is the predefined accuracy.
-%Также алгоритм останавливается, если превышено заданное при запуске число испытаний $K_{max}$.
 
+%Russian
+Также алгоритм останавливается, если превышено заданное при запуске число испытаний $K_{max}$.
 The considered variant of the parallelization is \textit{synchronous}, when the transition to the next iteration is performed after full termination of the current one, i.e., after completing the last trial of the current iteration.
-% Конкретная реализация данной схемы распараллеливания будет зависеть от архитектуры вычислительной системы, а также от требований к программному обеспечению, необходимому для вычисления значений целевой функции задачи.
-В данном исследовании проведение испытаний требовало решения жесткой системы ODE (для ее решения был использован метод Radau IIA из бибилиотеки SciPy языка Python).
-Поэтому распараллеливание было ораганизовано на CPU с использованием технологии MPI.
-В одном (master) процессе работал алгоритм глобального поиска; остальные (slave) процессы проводили параллельные испытания в рамках выполнения шага 5 алгорита.
-Объем передаваемых данных между процессами будет небольшим: требуется лишь разослать координаты точек испытаний и затем собрать вычисленные значения целевой функции в этих точках.
-Одновременно с этим процесс проведения испытания достаточно трудоемкий (время проведения одного испытания занимает не менее 1 сек.), что значительно превышает время передачи данных между процессами.
+Конкретная реализация данной схемы распараллеливания будет зависеть от архитектуры вычислительной системы, а также от требований к программному обеспечению, необходимому для вычисления значений целевой функции задачи.
+В данном исследовании проведение испытаний требовало решения жесткой системы ODE (для ее решения был использован метод Radau IIA).
+Поэтому распараллеливание было организовано на CPU с использованием технологии MPI.
+В одном (master) процессе работал алгоритм глобального поиска; остальные (slave) процессы проводили параллельные испытания в рамках выполнения шага 5 алгоритма.
+Объем передаваемых данных между процессами был небольшим: требовалось лишь разослать координаты точек испытаний и затем собрать вычисленные значения целевой функции в этих точках.
+Одновременно с этим время проведения одного испытания составляло не менее 1 сек., что значительно превышает время передачи данных между процессами.
+Тем самым, накладные расходы на передачу данных не оказывали существенного влияния на эффективность распараллеливания.
 
@@
 
-The numerical experiments were carried out using the Lobachevsky supercomputer (Lobachevsky State University of Nizhny Novgorod). 
-%Указать актуальную конфигурацию
-%Each supercomputer node included two quad-core processors Intel Xeon X5570, two NVIDIA Tesla X2070 and 12 Gb RAM.
-%Указать актуальное программное обеспечение
-%To build the program for running on the Lobachevsky supercomputer, the GCC 4.3.0 compiler, CUDA 6.5 and Intel MPI 2017 were used.
-%Для вычисления значений целевой функции задачи использовался Python .
+Numerical experiments were performed on the Lobachevsky supercomputer of the University of Nizhni Novgorod (operating system -- CentOS 7.2, control system -- SLURM).
+One supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM.
+The CPU is 8-core (i.e.
+a total of 16 CPU cores are available on the node).
+The numerical experiments were performed using the Globalizer system \cite{globalizerSystem}.
+To build the system for running on the Lobachevsky supercomputer, the GCC 5.5.0 compiler and Intel MPI 2017 were used.
 
-%Parallel global search algorithm запускался со следующими параметрами: параметр надежности $r=3$, точность поиска $\epsilon = 0.01$, плотность построения развертки $m=10$, максимальное число испытаний $K_{max}=1000$.
-После завершения работы parallel global search algorithm найденное решение уточнялось с помощью локального поиска, использовался Hooke--Jeeves method \cite{HookJeeves}.
+%Russian
+Для решения жесткой системы уравнений, возникающей при вычислении значений целевой функции, использовался метод Radau из библиотеки SciPy, Python 3.9.
+Global search algorithm запускался со следующими параметрами: параметр надежности $r=5$, точность поиска $\epsilon = 0.01$, плотность построения evolvent $m=10$, максимальное число испытаний $K_{max}=5000$.
+После завершения работы алгоритма найденное решение уточнялось с помощью локального поиска, использовался Hooke--Jeeves method \cite{HookJeeves}.
 Точность локального поиска составляла $\epsilon = 0.001$.
+Для оценки эффективности распараллеливания global search algorithm запускался в последовательном и паралелльном режимах, максимальное число задействованных вычислительных узлов составляло 20, соответственно, максимальное число задействованных процессоров было 40.
+На Fig.~\ref{fig_time},\ref{fig_speedup} приведены графики, отражающие время и ускорение параллельного алгортима в зависимости от числа процессоров; время работы приведено в часах.
 
-%Далее - описать параллельный запуск и его результаты, например, в виде таблицы с ускорением
+\begin{figure}[ht]
+\center{\includegraphics[width=0.8\linewidth]{time.pdf}}
+\caption{The parallrl algorithm running time}
+\label{fig_time}
+\end{figure}
 
-%Затем - проинтерпретировать результаты с содержателной точки зрения
+\begin{figure}[ht]
+\center{\includegraphics[width=0.8\linewidth]{speedup.pdf}}
+\caption{Speedup of the parallel algorithm}
+\label{fig_speedup}
+\end{figure}
 
-\section{Conclusion}
+\section{Conclusion }
 
-%планы дальнейших работ - реализация метода решения системы
+%Russian
+В планы дальнейших работ входит также и повышение эффективности распараллеливания используемого метода.
+Относительно невысокие показатели ускорения при решении данной задачи объясняются двумя обстоятельствами.
+Во-первых, при реализации параллельного алгоритма была применена синхронная схема распараллеливания.
+Управляющий процесс (в котором работал метод оптимизации) при выполнении  шага 5 алгоритма ожидал окончания проведения всех испытаний, проводимых подчиненными процессами.
+Тогда как в рассматриваемой задаче время проведения одного испытания (как было экспериментально установлено) варьировалось от 1 до 40 секунд в зависимости от конкретных значений параметров.
+Возможный путь к повышению эффективности распаралелливания - реализация асинхронной схемы проведения испытаний на шаге 5 алгоритма.
+В этом случае результаты испытания, полученные в одном из подчиненных процессов, могут быть отправлены для обработки в управляющий процесс немедленно, без ожидания окончания других испытаний.
+После чего подчиненный процесс получает от управляющего процесса новую точку, в которой проводит очередное испытание.
+Асинхронная схема распараллеливания, теоретически, способна обеспечить полную загрузку задействованных вычислительных ресурсов даже в том случае, когда время выполнения испытаний различно в разных точках области поиска.
+
+Во-вторых, при масштабном распараллеливании "узким местом" становится локальный поиск, который выполняется для уточнения решения после завершения фазы глобального поиска.
+Локальный поиск требует незначительного (по сравнению с глобальным) числа trials, но эти испытания выполняются последовательно.
+И при паралелльном запуске это незначительное число последовательных испытаний становится сопоставимым с числом параллельных итераций глобального поиска.
+В дальнейшем планируется использовать параллельную реализацию метода локального поиска на завершающем этапе работы, что способно повысить эффективность распараллеливания в целом.
 

### 2021-04-15 2021/RSD_Ufa/RSD2021.tex
Translate our part

@@
 
-%Russian
-Метод решения оптимизационных задач подобного типа должен учитывать следующую специфику.
+%Метод решения оптимизационных задач подобного типа должен учитывать следующую специфику.
 Во-первых, целевая функция задачи является "черным ящиком", т.к.
@@
 С одной стороны, вычисление градиента в данном случае будет чрезмерно дорогой операцией, а с другой стороны - мультистартовые методы слабо применимы для существенно многоэкстремальных задач, какими и являются обратные задачи химической кинетики.
-
 Перечисленные особенности inverse problem of chemical kinetics учитываются в разработанном в University of Nizhni Novgorod параллельном алгоритме глобальной оптимизации.
@@
 В данной статье представлены результаты, полученные при использовании parallel global search algorithm для решения inverse problem of chemical kinetics.
+
+A method for solving the optimization problems of such type should account for the following aspects.
+First, the objective function of the problem is a ``black box'' since it is defined not by a formula but in the form of some algorithm for computing its values.
+Second, computing even a single value of the objective function is time-consuming operation since it requires numerical modeling.
+The aspects listed above don't allow employing the multistart methods based on the ideas of multiple gradient descents for solving the inverse problems.
+On one hand, computing the gradients would be too expensive operation in this case.
+On the other hand, the multistart methods are hardly applicable for the essentially multiextremal problems, which the inverse problems of chemical kinetics belong to.
+
+The features of inverse problems of chemical kinetics listed above were taken into account in the parallel global search algorithm developed in University of Nizhni Novgorod.
+The information-statistical approach described in \cite{Strongin2000} in details makes a theoretical basis for this algorithm.
+The developed algorithm is based on a reduction of the initial multidimensional problem to an equivalent one-dimensional problem followed by solving the latter by efficient optimization methods for the functions of single variable.
+The issues of the parallelization of the algorithm on various architectures were considered in \cite{Barkalov2016}.
+In the present paper, the results obtained with the use of the parallel global search algorithm for solving the inverse problem of chemical kinetics are presented.
 
@@
 The algorithm stops if condition $\Delta_{t_j}<\epsilon$ is satisfied for at least one index $t_j, 1 \leq j \leq p$ ; here $\epsilon>0$ is the predefined accuracy.
+Also, the algorithm stops if the maximum allowed number of trials $K_{max}$ preset prior to running is exhausted.
 
 %Russian
-Также алгоритм останавливается, если превышено заданное при запуске число испытаний $K_{max}$.
 The considered variant of the parallelization is \textit{synchronous}, when the transition to the next iteration is performed after full termination of the current one, i.e., after completing the last trial of the current iteration.
-Конкретная реализация данной схемы распараллеливания будет зависеть от архитектуры вычислительной системы, а также от требований к программному обеспечению, необходимому для вычисления значений целевой функции задачи.
+Particular implementation of this parallelization scheme will depend on the computer system architecture as well as on the requirements to the software necessary for computing the objective function values.
+In the present study, the conducting of trials required solving a stiff ODE system (for solving this one, Radau IIA method was used).
+Therefore, the parallelization was organized on CPU using MPI technology.
+The global search algorithm was run on a single (master) processor.
+The rest (slave) processors conducted parallel trials within the Step 5 of the algorithm.
+The amount of data transferred between the processes was small: it was only required to send out the coordinates of trial points and then to collect the computed values of the objective function at these points.
+At the same time, the time of conducting a single trial was not less 1 sec that considerably exceeded the time of data transfer between the processes.
+This way, the overhead costs for the data transfer didn't essentially affect the parallelization efficiency.
+
+%Конкретная реализация данной схемы распараллеливания будет зависеть от архитектуры вычислительной системы, а также от требований к программному обеспечению, необходимому для вычисления значений целевой функции задачи.
 В данном исследовании проведение испытаний требовало решения жесткой системы ODE (для ее решения был использован метод Radau IIA).
@@
 To build the system for running on the Lobachevsky supercomputer, the GCC 5.5.0 compiler and Intel MPI 2017 were used.
+To solve the stiff system of equations arising in computing the objective function values, Radau method from SciPy library from Python 3.9 was used.
 
-%Russian
-Для решения жесткой системы уравнений, возникающей при вычислении значений целевой функции, использовался метод Radau из библиотеки SciPy, Python 3.9.
+%Для решения жесткой системы уравнений, возникающей при вычислении значений целевой функции, использовался метод Radau из библиотеки SciPy, Python 3.9.
 Global search algorithm запускался со следующими параметрами: параметр надежности $r=5$, точность поиска $\epsilon = 0.01$, плотность построения evolvent $m=10$, максимальное число испытаний $K_{max}=5000$.
@@
 На Fig.~\ref{fig_time},\ref{fig_speedup} приведены графики, отражающие время и ускорение параллельного алгортима в зависимости от числа процессоров; время работы приведено в часах.
+
+Global search algorithm was run with the following parameters: the reliability parameter $r=5$, the search precision $\epsilon = 0.01$, the evolvent density $m=10$, the maximum allowed number of trials $K_{max}=5000$.
+Upon completing the execution of the algorithm, the solution found was refined with a local search using Hooke--Jeeves method \cite{HookJeeves}.
+The local search precision was $\epsilon = 0.001$.
+To evaluate the parallelization efficiency, the global search algorithm was run in the sequential regime and in the parallel one, the maximum number of computer nodes employed was 20.
+Correspondingly, the maximum number of processors employed was 40.
+In Fig.~\ref{fig_time} and Fig.~\ref{fig_speedup} the plots reflecting the running time and the speedup of the parallel algorithm {\em vs} the number of processors employed, respectively are presented (the time is in hours).
 
@@
 
-\section{Conclusion }
+\section{Conclusions and Future Work}
 
-%Russian
-В планы дальнейших работ входит также и повышение эффективности распараллеливания используемого метода.
+%В планы дальнейших работ входит также и повышение эффективности распараллеливания используемого метода.
 Относительно невысокие показатели ускорения при решении данной задачи объясняются двумя обстоятельствами.
@@
 Асинхронная схема распараллеливания, теоретически, способна обеспечить полную загрузку задействованных вычислительных ресурсов даже в том случае, когда время выполнения испытаний различно в разных точках области поиска.
-
 Во-вторых, при масштабном распараллеливании "узким местом" становится локальный поиск, который выполняется для уточнения решения после завершения фазы глобального поиска.
@@
 В дальнейшем планируется использовать параллельную реализацию метода локального поиска на завершающем этапе работы, что способно повысить эффективность распараллеливания в целом.
+
+The plans of further work include also the improvement of the parallelization efficiency for the global optimization method used.
+Relatively low indicators of speedup in the solving of the present problem can be explained by two factors.
+First, a synchronous parallelization scheme was applied in the implementation of the parallel algorithm.
+The master process (which the optimization method worked in) was waiting for the finish of execution of all trials performed by the slave processes when executing Step 5 of the algorithm.
+However, the time of executing  a single trial in the problem considered varied from 1 up to 40 seconds subject to particular values of parameters (as it was found experimentally).
+The implementation of the asynchronous scheme of executing the trials at Step 5 of the algorithm seems to be a promising way of improving the parallelization efficiency.
+In this case, the trial results obtained by a slave process may be sent to the master process for processing immediately without waiting for the competing of other trials.
+Afterwards, the slave process receives a new point of the next trial form the master one.
+Theoretically, the asynchronous parallelization scheme can provide a full load of the computer resources employed even if the trial execution times are different at different points of the search domain.
+
+Second, the local search, which is performed to refine the solution after completing the global search phase becomes a ``bottleneck'' at large-scale parallelization.
+The local search requires insufficient number of trials (as compared to the global one), but these trials should be performed sequentially.
+And at the parallel run this insufficient number of sequential trials becomes comparable to the number of the parallel global search iterations.
+In future, we plan to use the parallel implementation of the local search method in the final stage of the work that can improve the overall parallelization efficiency.
 

### 2021-04-15 2021/RSD_Ufa/RSD2021.tex
new figures from Ufa

@@
 
-\usepackage[russian]{babel}
+%\usepackage[russian]{babel}
 
@@
 %
-\titlerunning{Adaptive Global Optimization}  % abbreviated title (for running head)
+\titlerunning{Parallel Global Search}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 To do this, the inverse problem of chemical kinetics is solved, which predicts the values of kinetic parameters based on laboratory data.
-From a mathematical point of view, the inverse of the chemical kinetics problem is a global optimization problem.
+From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem.
 A parallel information-statistical global search algorithm was used to solve it.
-The use of the parallel algorithm has reduced significantly the search time for the optimum.
+The use of the parallel algorithm has significantly reduced the search time to find the optimum.
 The found optimal parameters of the model made it possible to adequately simulate the process of pre-reforming propane on a Ni-catalyst.
 
-\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing }
+\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Chemical kinetics $\cdot$ Inverse problems }
 \end{abstract}
@@
 One of the areas of the Laboratory of Mathematical Chemistry of the Institute of Petrochemistry and Catalysis of the UFIC RAS is the study of chemical reactions by mathematical modeling methods.
-At the moment, a large number of complex chemical processes have been studied-the reaction of cycloalumination of alkenes with triethylaluminium into alumacyclopentanes, the reaction of hydroalumination of olefins with organoaluminum compounds, the reaction of alcohols with dimethyl carbonate, etc.
-The mathematical description of chemical reactions is necessary for the construction of mathematical models of chemical kinetics, which are called kinetic models.
+At the moment, a large number of complex chemical processes have been studied, among  them, the reaction of cycloalumination of alkenes with triethylaluminium into alumacyclopentanes, the reaction of hydroalumination of olefins with organoaluminum compounds, the reaction of alcohols with dimethyl carbonate, etc.
+The mathematical description of chemical reactions is necessary for the development the mathematical models of chemical kinetics, which are called kinetic models.
 The kinetic model is the basis for mathematical modeling of chemical reactions and the initial level for mathematical modeling of chemical reactors.
 When developing a kinetic model of a multi-stage chemical reaction, the inverse problem of chemical kinetics (CK) is solved.
-The inverse problem of HC is generally called the process of selecting the mechanism of a complex process and determining its quantitative characteristics - the rate constants of single-stage reactions based on experimental data on the dynamics of the process.
+The inverse problem of CK is generally called the process of selecting the mechanism of a complex process and determining its quantitative characteristics -- the rate constants of reactions based on experimental data about the dynamics of the process.
 From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem, so the problem of choosing an optimization method is relevant.
@@
 The developed algorithm is based on a reduction of the initial multidimensional problem to an equivalent one-dimensional problem followed by solving the latter by efficient optimization methods for the functions of single variable.
-The issues of the parallelization of the algorithm on various architectures were considered in \cite{Barkalov2016}.
+The issues of the parallelization of the algorithm on various architectures were considered in \cite{Barkalov2016,Strongin2018,Barkalov2020}.
 In the present paper, the results obtained with the use of the parallel global search algorithm for solving the inverse problem of chemical kinetics are presented.
@@
 \end{gather}
-
 where the rates of reactions \eqref{reac_ref} -- \eqref{reac_met} are expressed according to the Langmuir-Hinshelwood model \cite{USKOV2019126741}:
@@
 \]
-
-where $W_{ref}$ and $W_{met}$ are the reaction rates; $E_{ref}$ and $E_{met}$ are the observed activation energies, J/mol; $k^0_{ref}$ and $k^0_{met}$ are the pre–exponential multipliers; $B$ is the constant parameter, $T$ is the temperature, K; $R$ is the universal gas constant, J(K$\cdot$ mol).
-The "$ref$"and "$met$" indexes refer to pre-reforming and methanation reactions, respectively.
+where $W_{ref}$ and $W_{met}$ are the reaction rates; $E_{ref}$ and $E_{met}$ are the observed activation energies, J/mol; $k^0_{ref}$ and $k^0_{met}$ are the pre–exponential multipliers; $B$ is the constant parameter, $T$ is the temperature, K; $R$ is the universal gas constant, J/(K$\cdot$ mol).
+The ``$ref$''and ``$met$'' indexes refer to pre-reforming and methanation reactions, respectively.
 $C_{C3H8}$ and $C_{H2}$ are concentrations of propane and hydrogen, mol/m$^3$; $m$ is order of the denominator, which varied from 0 to 2; $K_{eq}$ is the equilibrium constant of $CO_2$ methanation; $P_{CH4}$, $P_{H2O}$, $P_{CO2}$, $P_{H2}$ are partial pressures of the corresponding substances, bar.
@@
 \end{equation}
-
 where $G$ is a mass flow of the mixture, kg/(m$^2 \cdot$ sec); $y_i$ is a mass fraction of the $i$-th component; $\nu_i$ is a stoichiometric coefficient of the $i$-th component; $m_i$ is a molar mass of the $i$-th component, kg/mol; $l$ is coordinate along the catalytic layer, m; $L$ is a length of the catalytic layer, m.
@@
 
-Systems of ordinary differential equations (ODES) describing chemical processes are often rigid due to the presence of fast-flowing and slow-flowing reaction stages.
-Therefore, the RADAU-IIA method was chosen as a method for solving the ODE system, which is suitable for solving rigid ODE systems.
+Systems of ordinary differential equations (ODES) describing chemical processes are often stiff due to the presence of fast-flowing and slow-flowing reactions.
+Therefore, the RADAU-IIA method was chosen as a method for solving the ODE system, which is suitable for solving stiff ODE systems.
 Let us consider in the general case the Cauchy problem of the form:
@@
 \end{equation}
-
 %где x в нашем случае это вектор концентраций веществ реакции размерности 5, а f это правая часть уравнения \eqref{mater_balance}.
 Тогда s-стадийный метод Рунге-Кутты для решения системы \eqref{ode} выписывается следующим образом:
-
-where $x$ in our case is the vector of concentrations of reaction substances of dimension 5, and $f$ is the right-hand side of the equation \eqref{mater_balance}.
+where $x$ is the vector of concentrations of reaction substances with dimension of 5 (the substances are $C_3H_8$, $CH_4$, $H_2O$, $H_2$, $CO_2$), and $f$ is the right-hand side of the equation \eqref{mater_balance}.
 Then the $s$-stage Runge-Kutta method for solving the \eqref{ode} system is written as follows:
@@
 \end{equation}
-
 %где $x_n$, $x_{n+1}$, $X_i$ -- приближенные решения системы \eqref{ode} в моменты времени $t_n$, $t_{n+1} = t_n + h$, $t_n + c_i h$.
 Коэффициенты  методов  Рунге-Кутты  обычно  записывают  в  виде таблицы Бутчера:
-
-where $x_n$, $x_{n+1}$, $X_i$ are approximate solutions of the \eqref{ode} system at time points $t_n$, $t_{n+1} = t_n + h$, $t_n + c_i h$.
-The coefficients of the Runge-Kutta methods are usually written in the form of a Butcher table:
+where $x_n$, $x_{n+1}$, $X_i$ are approximate solutions of the system \eqref{ode} at time points $t_n$, $t_{n+1} = t_n + h$, $t_n + c_i h$.
+The coefficients of the Runge-Kutta methods are usually written in the form of a Butcher tableau:
 
@@
 
-The Butcher table contains the vectors $b$ and $c$, consisting of the coefficients $b_i$ and $c_i$, respectively, and the matrix $A$ of dimension $s \times s$, including the coefficients $a_{ij}$.
-
-Fifth-order Rado-IIA method (Radau 5):
+The Butcher tableau contains the vectors $b$ and $c$, consisting of the coefficients $b_i$ and $c_i$, respectively, and the matrix $A$ of dimension $s \times s$, including the coefficients $a_{ij}$.
+The coefficients of fifth-order Rado-IIA method (Radau 5) are presented below:
 
@@
 
+The solution of the system of ordinary differential equations is a vector of the calculated concentrations of the reaction components, which are then compared with the experimental data.
 Determining the kinetic parameters of reaction stages by comparing calculated values of substance concentrations and experimental results is an inverse problem of chemical kinetics.
 The mathematical problem is to minimize the functional of the deviation between calculated and experimental values.
-The functional of minimizations determined as the sum of absolute deviations between calculated and experimental concentrations:
+The objective function determined as the sum of absolute deviations between calculated and experimental concentrations:
 
@@
 \end{equation}
-
 where $x_{ij}^{calc}$ and $x_{ij}^{exp}$ are calculated and experimental values of component concentrations; $M$ is the number of measuring points; $N$ is the number of substances involved in the reaction.
@@
 \section{Numerical Experiments}\label{Sec_Exp}
+%Таким образом, в результате решения обратной задачи химической кинетики параллельным методом глобального поиска были рассчитаны кинетические параметры процесса предриформинга пропана на Ni катализаторе (Table \ref{tab:1}).
+Для интерпретации полученных результатов проводится решение прямой задачи химической кинетики, то есть расчет концентраций компонентов реакции и сравнение их с экспериментальными данными.
+На рисунках НОМЕР a, b показано сравнение рассчитанных концентраций с экспериментальными при разных условиях: (a) GHSV 4000 h$^{-1}$, (b) GHSV 12000 h$^{-1}$.
+Результаты расчетов качественно согласуются с экспериментом.
+Ошибка отклонения составляет не более ПОСЧИТАТЬ \%.
+Thus, as a result of solving the inverse problem of chemical kinetics by the parallel global search method, the kinetic parameters of the propane pre-reforming process on a Ni catalyst were calculated (Table \ref{tab:1}).
+To interpret the results obtained, the solution of the direct problem of chemical kinetics was carried out, namely, the concentrations of the reaction components were calculated and compared with the experimental data.
+Fig.~\ref{fig_kin_curves} shows a comparison of the calculated concentrations with the experimental ones under different conditions: (a) GHSV=4000 h$^{-1}$, (b) GHSV=12000 h$^{-1}$.
+The results of the calculations are qualitatively consistent with the experiment.
+
+\begin{figure}[ht]
+	\begin{minipage}{0.5\linewidth}
+		\center{\includegraphics[width=1.0\linewidth]{kin_curve1.png} \\ (a)}
+	\end{minipage}
+	\begin{minipage}{0.5\linewidth}
+		\center{\includegraphics[width=1.0\linewidth]{kin_curve2.png} \\ (b)}
+	\end{minipage}
+	\caption{Temperature dependences of the output concentrations of propane C3H8, methane CH4, hydrogen H2 and CO2 in the process of propane pre-reforming.
+Experimental conditions: 220-380 $^0$C, (a) GHSV = 4000 h$^{-1}$, (b) GHSV = 12000 h$^{-1}$, 1 bar pressure.
+Concentrations of the gas components on the figure are given on the dry basis.
+Points are experimental concentrations (``exp''-index), lines are simulated concentrations (``calc''-index).}
+	\label{fig_kin_curves}
+\end{figure}
+
+\begin{table}[ht]
+	\caption{Calculated kinetic parameters}
+	\label{tab:1}
+	\center
+	\begin{tabular}{c|c|c|c|c|c}
+		\hline
+		%\noalign{\smallskip}
+		$E_{ref}$, kJ/mol & $E_{met}$, kJ/mol & $k_{ref}$, (mole$\cdot$m$^{-3}$)m$\cdot$s$^{-1}$ & $k_{met}$, s$^{-1}$ & $m$ & $B$ \\
+		%\noalign{\smallskip} 
+		\hline 
+		%\noalign{\smallskip}
+		99.23 \;&	37.08 \; & $2.4 \cdot 10^{10}$ \;  & $1.93 \cdot 10^5$ \; & 0.82 & 4.8  \\
+		%\noalign{\smallskip}
+		\hline
+	\end{tabular}
+\end{table}
+%Для сравнения с предыдущими расчетами был построен график арренисовской зависимости (рисунок НОМЕР), на котором представлены зависимости десятичного логарифма от констант скоростей реакций \eqref{reac_ref} и \eqref{reac_met} от обратной температуры.
+Обозначения "GA", "GSA" и "PGSA" означают  решения, полученные различными алгоритмами оптимизации, соотвественно, генетическим алгоритмом, алгоритмом гравитационного поиска и параллельным алгоритмом глобального поиска.
+
+For comparison with the previous calculations, an Arrhenis plot was constructed (Fig.~\ref{fig_arren}), which shows the dependence of the decimal logarithm on the reaction rate constants \eqref{reac_ref} and \eqref{reac_met} on the inverse temperature.
+The designations ``GA'', ``GSA'' and ``PGSA'' mean solutions obtained by various optimization algorithms, the genetic algorithm, the gravity search algorithm, and the parallel global search algorithm, respectively.
+The blue lines refer to the pre-reforming reaction \eqref{reac_ref}, and the yellow lines refer to the methanation reaction~\eqref{reac_met}.
+Note that the lines of the same color are parallel to each other, which means that the activation energies of the reaction calculated by different methods have approximately the same values, which indicates the adequacy of the calculations.
+
+\begin{figure}[ht]
+	\center{\includegraphics[width=0.8\linewidth]{arren.png}}
+	\caption{Comparison of the obtained solution with the other algorithms found on the Arrhenius plot}
+	\label{fig_arren}
+\end{figure}
 
@@
 \center{\includegraphics[width=0.8\linewidth]{time.pdf}}
-\caption{The parallrl algorithm running time}
+\caption{The parallel algorithm running time}
 \label{fig_time}
@@
 \section{Conclusions and Future Work}
+
+%В будущем планируется моделирование других химических процессов, а именно, первая задача это моделирование процесса сернокислотного алкилирования изобутана олефинами, для которого на данный момент составлена схема химических превращений, проведен литературный обзор процесса и найдены экспериментальные данные.
+Математическая модель реакции сернокислотного алкилирования изобутана олефинами представляет собой задачу Коши, на данный момент решена прямая задача химической кинетики, далее будет решаться обратная задача предложенными параллельными методами.
+Вторая задача это моделирование процесса синтеза метил-трет бутилового эфира.
+Химизм данного процесса известен, задача иемют небольшую размерность, однако моделирование реактора трудоемко и представляет собой решение уравнение в частных производных второго порядка.
+На данный момент расписана разностная схема, реализован метод прогонки, следующим этапом будет решение прямой задачи химической кинетики, а также предсказание кинетических параметров процесса предложенными параллельными методами.
+
+In the future, it is planned to simulate other chemical processes, namely, the first task is to simulate the process of sulfuric acid alkylation of isobutane with olefins, for which a scheme of chemical transformations has been compiled, a literature review of the process has been conducted, and experimental data have been found.
+The mathematical model of the reaction of sulfuric acid alkylation of isobutane by olefins is a Cauchy problem, at the moment the direct problem of chemical kinetics is solved, then the inverse problem will be solved by the proposed parallel methods.
+The second task is to simulate the synthesis of methyl-tert-butyl ether.
+The chemistry of this process is known, and the problem has a small dimension, but the simulation of the reactor is time-consuming and represents the solution of a second-order partial differential equation.
+At the moment, the difference scheme is formed, the run-through method is implemented, the next stage will be the solution of the direct problem of chemical kinetics, as well as the prediction of the kinetic parameters of the process by the proposed parallel methods.
 
@@
 
-\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204.
+\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204 and by RFBR, project No.\,19-37-60014.
 

### 2021-05-25 2021/RSD_Ufa/RSD2021.tex
add three new references

@@
 First, the objective function of the problem is a ``black box'' since it is defined not by a formula but in the form of some algorithm for computing its values.
-Second, computing even a single value of the objective function is time-consuming operation since it requires numerical modeling.
+Second, computing even a single value of the objective function is time-consuming operation since it requires numerical modeling \cite{Modorskii2016}.
 The aspects listed above don't allow employing the multistart methods based on the ideas of multiple gradient descents for solving the inverse problems.
@@
 
-The parallel global search algorithm described in this section is based on the idea of reducing the dimension with the use of the Peano curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously map the unit interval $[0,1]$ onto the $N$-dimensional cube $D$ from (\ref{D}).
+Many state-of-the-art global optimization algorithms use the idea of reducing the dimensionality for solving multidimensional problems; see, for example, the diagonal partitions method \cite{Sergeyev2006} or the simplicial partitions method \cite{Zilinskas2008}.
+The parallel global search algorithm described in this section uses the Peano curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously map the unit interval $[0,1]$ onto the $N$-dimensional cube $D$ from (\ref{D}).
 By using this kind of mapping, it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem

### 2021-05-30 2021/RSD_Ufa/RSD2021.tex
RSD2021 final

@@
 
-One of the areas of the Laboratory of Mathematical Chemistry of the Institute of Petrochemistry and Catalysis of the UFIC RAS is the study of chemical reactions by mathematical modeling methods.
+One of the areas of the Laboratory of Mathematical Chemistry of the Institute of Petrochemistry and Catalysis of the UFIC RAS is the study of chemical reactions by mathematical modeling methods \cite{Zaynullin2020, Zainullin2019, Gubaidullin2014}.
 At the moment, a large number of complex chemical processes have been studied, among  them, the reaction of cycloalumination of alkenes with triethylaluminium into alumacyclopentanes, the reaction of hydroalumination of olefins with organoaluminum compounds, the reaction of alcohols with dimethyl carbonate, etc.
-The mathematical description of chemical reactions is necessary for the development the mathematical models of chemical kinetics, which are called kinetic models.
+The mathematical description of chemical reactions is necessary for the development the mathematical models of chemical kinetics, which are called kinetic models \cite{Koledina2019}.
 The kinetic model is the basis for mathematical modeling of chemical reactions and the initial level for mathematical modeling of chemical reactors.
@@
 The inverse problem of CK is generally called the process of selecting the mechanism of a complex process and determining its quantitative characteristics -- the rate constants of reactions based on experimental data about the dynamics of the process.
-From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem, so the problem of choosing an optimization method is relevant.
+From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem, so the problem of choosing an optimization method is relevant \cite{Enikeeva2021, Enikeeva2020}.
 
@@
 
-The process under study is propane pre-reforming into methane-rich gas over Ni catalyst, which is an industrially important chemical process \cite{USKOV2019126741, Uskov_en}.
+The process under study is propane pre-reforming into methane-rich gas over Ni catalyst, which is an industrially important chemical process \cite{Uskov2019, Uskov2020}.
 Pre-reforming of propane was studied over industrial nickel-chromium catalyst at pressure of 1 and 5 bar in the temperature range of 220 -- 380 $^0$C and at flow rates of 4000 and 12000 h$^{-1}$.
-The experimental data on propane pre-reforming was acquired from our previous work \cite{Uskov_en}.
-The reaction scheme consists of two reactions: propane steam conversion and CO$_2$ methanation \cite{Uskov_cat}:
+The experimental data on propane pre-reforming was acquired from our previous work \cite{Uskov2020}.
+The reaction scheme consists of two reactions: propane steam conversion and CO$_2$ methanation \cite{Uskov2017}:
 \begin{gather}
@@
 \end{gather}
-where the rates of reactions \eqref{reac_ref} -- \eqref{reac_met} are expressed according to the Langmuir-Hinshelwood model \cite{USKOV2019126741}:
+where the rates of reactions \eqref{reac_ref} -- \eqref{reac_met} are expressed according to the Langmuir-Hinshelwood model \cite{Uskov2019}:
 
@@
 The mathematical model of chemical kinetics problems is a system of differential equations that describes the variations in substance concentrations over time according to the rates of reaction stages.
-The system of differential equations is a Cauchy problem containing the initial data \cite{ENIKEEV2020123131, Chainikova}.
+The system of differential equations is a Cauchy problem containing the initial data \cite{Enikeev2020, Chainikova2017}.
 The numerical solving of such a system of equations is a direct problem of chemical kinetics.

### 2021-06-10 2021/ICCS_BarkalovUsova/iccs.tex
Include one more comment in Russian

@@
 
-\usepackage{hyperref}
+%\usepackage{hyperref}
 
@@
 %%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
-%\usepackage[utf8x]{inputenc}
-%\usepackage[english,russian]{babel}
+\usepackage{inputenc}
+\usepackage[english,russian]{babel}
 %\usepackage{cmap}
@@
 \end{equation}
-One-dimensional optimization problems play an important role, because many approaches to solving multidimensional optimization problems are, in one way or another, based on reducing the solution of the original multidimensional problem to solving a series of related one-dimensional optimization problems (see, for example, popular approaches in \cite{PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
+
+Задачи глобальной оптмизации, которые характеризуются наличием многих локальных экстремумов, возникают во многих областях приложений (see, e.g., \cite{Cavoretto2021,Kvasov2015,Kvasov2013,Modorskii2016}).
+Как правило, вычисление даже одного значения целевой функции в таких задачах характеризуется значительным временем, т.к.
+требует трудоемкого численного моделирования.
+В этом случае принято говорить о expensive black-box optimization problems.
+Задачи указанного класса чрезвычайно сложны и являются объектом интенсивных исследований.
+
+Следует отметить важную роль одномерных задач глобальной оптимизации. Во-первых, как уже отмечалось, они возникают во многих приложениях (see, e.g., \cite{Gillard2017,Sergeyev2020}). А во-вторых,  
+ many approaches to solving multidimensional optimization problems are, in one way or another, based on reducing the solution of the original multidimensional problem to solving a series of related one-dimensional optimization problems (see, for example, popular approaches in \cite{Pinter1996,PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
 
@@
 where the constant $L$ is unknown a priori. 
-
 This feature of the objective function is commonly used in many approaches to the development of algorithms for multiextremal optimization \cite{Evtushenko2009,Evtushenko2013,Jones2009,Paulavicius2011}.

### 2021-06-11 2021/ICCS_BarkalovUsova/iccs.tex
Add ORCID

--- 2020-12-16 2021/ICCS_BarkalovUsova/iccs.tex
@@
 %
-\title{An algorithm for searching the global minimum of univariate discontinuous functions\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.
-0729-2020-0055, and by the Research and Education Mathematical Center, project no.
-075-02-2020-1483/1.}}
+\title{An algorithm for searching the global minimum of univariate discontinuous functions\thanks{This study was supported by the Russian Science Foundation, project No.\,21-11-00204.}}
 %
@@
 %
-\author{Konstantin Barkalov %\orcidID{0000-0001-5273-2471}  
-\and Marina Usova %\orcidID{0000-0002-8736-0652}
+\author{Konstantin Barkalov \orcidID{0000-0001-5273-2471}  
+\and Marina Usova \orcidID{0000-0002-0722-6884}
 }
@@
 
-\keywords{Global optimization \and Multiextremal functions \and 
-Discontinuous functions \and Algorithms' comparison }
+\keywords{Global optimization \and Multiextremal functions \and Discontinuous functions \and Algorithms' comparison }
 
@@
 
-\begin{figure}[!ht]
+\begin{figure}[ht]
 	\begin{center}

### 2021-06-11 2021/ICNAAM/2021_ICNAAM.tex

\documentclass{aip-cp}

\usepackage[numbers]{natbib}
\usepackage{rotating}
\usepackage{graphicx}

% Document starts
\begin{document}

% Title portion
\title{The Title Goes Here with Each Initial Letter Capitalized}

\author[aff1]{Konstantin Barkalov\corref{cor1}}
\author[aff1]{Ilya Lebedev}
\author[aff2]{Daniil Ryazanov}
\author[aff2]{Sergey Strijak}

\affil[aff1]{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia}
\affil[aff2]{Ivannikov Institute for System Programming of the RAS, Moscow, Russia}
\corresp[cor1]{Corresponding author: konstantin.barkalov@itmm.unn.ru}

\maketitle

\begin{abstract}
The AIP Proceedings article template has many predefined paragraph styles for you to use/apply as you write your paper. To format your abstract, use the \LaTeX template style: {\itshape Abstract.} Each paper must include an abstract. Begin the abstract with the word ``Abstract'' followed by a period in bold font, and then continue with a normal 9 point font.
\end{abstract}

% Head 1
\section{INTRODUCTION}
Do not abbreviate Figure, Equation, etc.; display items are always singular, i.e., Figure 1 and 2. Equations are always singular, i.e., Equation 1 and 2, and should be inserted using the Equation Editor, not as graphics, in the main text.  Display items and captions should be inserted after the reference section. Please do not use footnotes in the text, additional information can be added to the reference list.

This is the paragraph spacing that occurs when you use the [ENTER] key.

\section{FIRST-ORDER HEADING}
This is an example of dummy text. This is an example of dummy text. This is an example of dummy text.
This is an example of dummy text. This is an example of dummy text. This is an example of dummy text.

This is the paragraph spacing that occurs when you use the [ENTER] key.

% enumerate
\noindent Numbered lists may also be included and should look like this:

\begin{enumerate}
\item This is an example of numbered listing.
\item This is an example of numbered listing.
\end{enumerate}

\noindent Bulleted lists may be included and should look like this:
% itemize

\begin{itemize}
\item This is an example of bulleted listing.
\item This is an example of bulleted listing. This is an example of bulleted listing.
\end{itemize}

\noindent This is an example of single line numbered equation.
\begin{equation}
\frac{d[F_1]}{d\omega_2} = SAm_2\cos\omega,\frac{d[F_1]}{d\omega_3}= SAm_2\cos\omega.
\end{equation}

This is an example of a multiline numbered equation.
% Numbered multiple equation
\begin{eqnarray}
p_{t_{10,1}}&=&\left(\frac{N_{cu}^2}{ N_c ^2}\right)\left(\frac{N_{ar}^2}{N_a^2}\right)\left(\frac{N_{ar}-1}{N_{ar}}\right),\\
p_{t_{10,2}}&=&\left(\frac{N_{cu}^2}{ N_c ^2}\right)\left(\frac{N_{ar}}{N_a^2}\right).
\end{eqnarray}
% Unnumbered equation
%% \begin{eqnarray*}
%% {\rm ((Multi\ Equation\ Unnumbered))}\\
%% {\rm ((Multi\ Equation\ Unnumbered))}\\
%% {\rm ((Multi\ Equation\ Unnumbered))}
%% \end{eqnarray*}
%% this is an example of dummy text.
% Unnumbered equation
%% \[
%% {\rm ((Single\ Equation\ Unnumbered))}
%% \]
For more on equations, please refer to the guide.

% Figure
%\begin{figure}[h]
  %\centerline{\includegraphics[width=150pt]{fig_1.pdf}}
  %\caption{To format a figure caption use the \LaTeX template style: Figure Caption. The text ``FIGURE 1,'' which labels the caption, should be bold and in upper case. If figures have more than one part, each part should be labeled (a), (b), etc. Using a table, as in the above example, helps you control the layout.}
%\end{figure}

Cite all figures in the text consecutively. The word ``Figure'' should be spelled out if it is the first word of the sentence and abbreviated as ``Fig.'' elsewhere in the text. Place the figures as close as possible to their first mention in the text at the top or bottom of the page with the figure caption positioned below the figure, all centered. Figures must be inserted in the text and may not follow the Reference section. Set figure captions in 9 point size, Times Roman font. Type the word ``FIGURE 1.'' in bold uppercase, followed by a period.\footnote{This is an example of a footnote.}

\subsection{Tables (Second Level Heading)}
Due to the wide range and complexity of tables, we simply offer an example for guidance. Please follow the style for Table \ref{tab:a} (and figure) captions.

% Table
\begin{table}[h]
\caption{To format a table caption, use the \LaTeX template style: Table Caption. The text ``{\bf TABLE 1,}'' which labels the caption, should be bold and all letters capitalized. Center this text above the table. Tables should have top and bottom rules, and a rule separating the column heads from the rest of the table only.}
\label{tab:a}
\tabcolsep7pt\begin{tabular}{lcccc}
\hline
  & \tch{1}{c}{b}{Single\\ outlet}  & \tch{1}{c}{b}{Small\\ multiple}  & \tch{1}{c}{b}{Large\\ multiple}  & \tch{1}{c}{b}{Total}   \\
\hline
1982 & 98 & 129 & 620    & 847\\
1987 & 138 & 176 & 1000  & 1314\\
1991 & 173 & 248 & 1230  & 1651\\
1998 & 200 & 300 & 1500  & 2000\\
\hline
\end{tabular}
\end{table}

\subsection{Font Embedding (Second Level Heading)}
As the author and creator of your article PDF, you have the most intimate knowledge of exactly what the PDF should display. We ask all authors to carefully check their article PDF prior to submission. Perform visual inspections to detect subtle font errors and ensure that all fonts are embedded. With the wide range of tools and software that authors use to create PDFs, and the number of devices and platforms that readers use to view/print them, font embedding by authors is not only ``nice-to-have'', it is essential.

\subsubsection{Why Should I Care About Font Embedding? (Third Level Heading)}
Embedding fonts into your PDF file is critically important for two reasons:
\begin{itemize}
\item Commercial printing companies are unable to print PDFs without the correct fonts embedded.
\item To ensure that your online article PDF file displays and prints correctly for everyone who wants to read your work.
\end{itemize}

Readers of scientific articles use an ever-increasing range of devices and applications to access, view, and print PDFs  from smart phones and tablets to desktop computers running any one of a number of operating systems. To ensure that readers of  your article can display and print it correctly, it is important for your article's PDF file to be truly portable. Your PDF file needs to be fully ``self-contained.''%\vfill\pagebreak

\section{FINAL KEY POINTS TO CONSIDER (FIRST LEVEL HEADING)}
Here are the main points you need to follow (the AIP Publishing author template packages contain comprehensive guidance):

\begin{itemize}
\item Write and prepare your article using the AIP Publishing template.
\item Create a PDF file of your paper (making sure to embed all fonts).
\item Send the following items to your conference organizer:
\begin{itemize}
\item PDF file of your paper
\item Signed Copyright Transfer Agreement
\item (If it applies) Copies of any permissions to re-use copyrighted materials in your article (e.g., figures from books/journals)
\end{itemize}
\end{itemize}

% Sections that will go in second font

% Acknowledgement
\section{ACKNOWLEDGMENTS}
This study was supported by the Russian Science Foundation, project No 21-11-00204.

\cite{Strongin2000,Kalyulin2017,Evtushenko2013,Jones2009,Zilinskas2010,Pinter1996,globalizerSystem,Barkalov2010,Strongin2020}

% References

%\nocite{*}
\bibliographystyle{aipnum-cp}%
\bibliography{bibliography}%

\end{document}

### 2021-06-15 2021/ICCS_BarkalovUsova/iccs.tex
Extend introduction

--- 2021-06-11 2021/ICCS_BarkalovUsova/iccs.tex
@@
 %%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
-%\usepackage[utf8x]{inputenc}
+%\usepackage{inputenc}
 %\usepackage[english,russian]{babel}
@@
 %
-\title{An algorithm for searching the global minimum of univariate discontinuous functions\thanks{This study was supported by the Russian Science Foundation, project No.\,21-11-00204.}}
+\title{A search algorithm for the global extremum \break of a discontinuous function\thanks{This study was supported by the Russian Science Foundation, project No.\,21-11-00204.}}
 %
@@
 \end{equation}
-One-dimensional optimization problems play an important role, because many approaches to solving multidimensional optimization problems are, in one way or another, based on reducing the solution of the original multidimensional problem to solving a series of related one-dimensional optimization problems (see, for example, popular approaches in \cite{PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
+
+Global optimization problems involving many local extrema arise in many areas of applications (see, e.g., \cite{Cavoretto2021,Kvasov2015,Kvasov2013,Modorskii2016}).
+As a rule, the calculation of even a single value of the objective function in such problems requires considerable time, since the analytical form of the function is not known and the calculation of its values involves time-consuming numerical modelling.
+In this case, we have to deal with expensive black-box optimization problems as they are often called.
+Problems of this class are extremely complicated and are the object of intensive research.
+
+One should note an important role of one-dimensional global optimization problems.
+Firstly, as has already been mentioned, they arise in many applications (see, e.g., \cite{Gillard2017,Sergeyev2020}).
+Secondly, many approaches to solving multidimensional optimization problems are, in one way or another, based on reducing the solution of the original multidimensional problem to solving a series of related one-dimensional optimization problems (see, for example, popular approaches in \cite{Pinter1996,PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
 

### 2021-06-15 2021/ICCS_BarkalovUsova/iccs.tex
correct references to the grants

@@
 %
-\title{A search algorithm for the global extremum \break of a discontinuous function\thanks{This study was supported by the Russian Science Foundation, project No.\,21-11-00204.}}
+\title{A search algorithm for the global extremum \break of a discontinuous function\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.
+0729-2020-0055, by the Research and Education Mathematical Сenter, project no.
+075-02-2020-1483/1, and by the Russian Foundation for Basic Research (project no.
+19-07-00242).} }
 %

### 2021-06-16 2021/ICCS_BarkalovUsova/iccs.tex
delete RFBR reference
update running head

@@
 \title{A search algorithm for the global extremum \break of a discontinuous function\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.
-0729-2020-0055, by the Research and Education Mathematical Сenter, project no.
-075-02-2020-1483/1, and by the Russian Foundation for Basic Research (project no.
-19-07-00242).} }
+0729-2020-0055, and by the Research and Education Mathematical Сenter, project no.
+075-02-2020-1483/1.}}
 %
-\titlerunning{An algorithm for searching the global minimum}
+\titlerunning{A search algorithm for the global extremum}
 % If the paper title is too long for the running head, you can set 

### 2021-06-22 2021/Entropy/entropy.tex
First sections of the Entropy paper

%  LaTeX support: latex@mdpi.com 
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[entropy,article,submit,moreauthors,pdftex]{Definitions/mdpi}

% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, analytica, animals, antibiotics, antibodies, antioxidants, appliedchem, applmech, applmicrobiol, applnano, applsci, arts, asi, atmosphere, atoms, audiolres, automation, axioms, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomechanics, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, brainsci, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coatings, colloids, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, crops, cryptography, crystals, curroncol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entropy, environments, environsciproc, epidemiologia, epigenomes, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forensicsci, forests, fractalfract, fuels, futureinternet, futuretransp, futurepharmacol, futurephys, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, land, languages, laws, life, liquids, literature, livers, logistics, lubricants, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, metabolites, metals, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, minerals, mining, modelling, molbank, molecules, mps, mti, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photochem, photonics, physchem, physics, physiolsci, plants, plasma, pollutants, polymers, polysaccharides, proceedings, processes, prosthesis, proteomes, psych, psychiatryint, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, systems, taxonomy, technologies, telecom, textiles, thermo, tourismhosp, toxics, toxins, transplantology, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wevj, women, world

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
\usepackage{inputenc}
\usepackage[T2A,T1]{fontenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}
%%%%

%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2021}
\copyrightyear{2020}
%\externaleditor{Academic Editor: Firstname Lastname} % For journal Automation, please change Academic Editor to "Communicated by"
\datereceived{} 
\dateaccepted{} 
\datepublished{} 
\hreflink{https://doi.org/} % If needed use \linebreak
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Title}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0001-5273-2471} % Add \orcidA{} behind the author's name
\newcommand{\orcidauthorB}{0000-0002-8736-0652} % Add \orcidB{} behind the author's name
\newcommand{\orcidauthorC}{0000-0002-4013-2329} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Konstantin Barkalov $^{1}$*\orcidA{}, Ilya Lebedev $^{1}$\orcidB{} and Victor Gergel $^{1}$\orcidC{}}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Konstantin Barkalov, Ilya Lebedev and Victor Gergel}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Barkalov, K.; Lebedev, I.; Gergel, V.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhni Novgorod, Russia; ilya.lebedev@itmm.unn.ru (I.L.); gergel@unn.ru (V.G.)}
%$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: konstantin.barkalov@itmm.unn.ru}
%; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}

% Current address and/or shared authorship
%\firstnote{Current address: Affiliation 3} 
%\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusion: indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia
%\encyclopediadef{Instead of the abstract}
%\entrylink{The Link to this entry published on the encyclopedia platform.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

%The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}. Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Histories, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences.

Успешное применение методов машинного обучения (ML) для решения широкого спектра задач приводит к созданию новых подходов к эффективному использованию идей ML в различных областях приложений.

Примером класса задач, в которых методы машинного обучения продемонстрировали свою эффективность, являются задачи выявления основных свойств исследуемых явлений (например, физических, экономических или социальных), которые характеризуются стохастической природой или наличием скрытых параметров \cite{Golovenkin2020,Gonoskov2019}.

Методы машинного обучения успешно используются и для решения сложных задач вычислительной математики, например, для simulation of dynamical systems \cite{Seleznev2019}, решения ordinary, partial or stochastic differential equations \cite{Lagaris1998,Blechschmidt2021,Xu2020} .

Одной из таких сложных задач вычислительной математики, к решению которой можно так или иначе применить методы машинного обучения, являются задачи глобальной оптимизации. 
В таких задачах, как правило, не представляется возможным найти решение аналитически и возникает необходимость построения численных методов для его поиска.

Проблема численного решения задач оптимизации сопряжена со значительными трудностями. Во многом они связаны с размерностью и видом целевой функции. При этом наиболее сложными являются задачи, в которых целевая функция является многоэкстремальной, недифференцируемой и, более того, заданной в форме черного ящика (т. е. в виде некоторой вычислительной процедуры, на вход которой подается аргумент, а выходом является соответствующее значение функции). Методы решения именно таких задач рассматриваются в данной статье.

Можно выделить несколько подходов к построению численных методов решения задач глобальной оптимизации.

Ряд алгоритмов основан на идее мультистарта: запуск локального поиска либо начиная с разных стартовых точек, либо с использованием разных параметров. Методы локальной оптимизации обладают большой скоростью сходимости. При этом одной из основных проблем в мультистартовых схемах является выбор начальных точек, которые бы соответствовали областям притяжения различных локальных решений. 
К решению данной проблемы могут быть успешно применены методы машинного обучения. 
Например, в \cite{RinnooyKan1987} для выбора перспективных стартовых точек использовались методы кластерного анализа. 
В \cite{Cassioli2012} область для запуска локального метода выделялась на основе классификации стартовых точке с помощью support vector machine.

Еще одним популярным классом методов, применимым к решению задач глобальной оптимизации, являются метаэвристические алгоритмы. 
Многие из них основаны на имитации процессов, протекающих в живой природе. Для настройки параметров таких алгоритмов также применимы методы машинного обучения. Например, в \cite{Jin2005} приведен обзор использования методов машинного обучения в эволюционных алгоритмах.

Следует отметить, что алгоритмы указанных двух классов не обеспечивают гарантированную сходимость к решению задачи и проигрывают детерминированным алгоритмам по качеству работы \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a certain set). Поэтому перспективным является использование именно детерминированных методов.

One of the efficient deterministic methods for solving multiextremal optimization problems is \textit{the information-statistical global search algorithm} \cite{Strongin2000}. This method initially proposed for solving unconstrained optimization problems was successfully generalized to the classes of optimization problems with non-convex constraints and multicriteria optimization problems. For different versions of the algorithm, parallelization methods taking into account the architecture of modern computing systems were also suggested \cite{Barkalov2016,globalizerSystem,Strongin2018}.

Для алгоритма глобального поиска в разное время было предложено несколько стратегий ускорения его работы (в терминах числа итераций, требующихся для решения задачи с заданной точностью). В данной работе предлагается новый подход к ускорению, основанный на выявления областей притяжения локальных минимумов с использованием методов машинного обучения. Выявление областей притяжения и запуск в этих областях локального поиска позволяет существенным образом сократить число испытаний, требующееся методу для достижения глобальной сходимости. Сказанное подтверждается результатами экспериментов, проведенных на серии из нескольких сотен тестовых задач.

\section{Problem Statement}

Будем рассматривать задачи глобальной оптимизации вида
\begin{eqnarray}\label{main_problem}
& \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
& D=\left\{y\in \text{R}^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
\end{eqnarray}
Задача \ref{main_problem} рассматривается в предположении, что целевая функция является многоэкстремальной, задана как ``черный ящик'', а вычисление ее значений связано с решением задачи численного моделирования и является трудоемкой операцией.

Для многих прикладных задач типичной является ситуация, когда ограниченное изменение вектора параметров $y$ вызывает ограниченное изменение значений $\varphi(y)$. Математической моделью, описывающей указанное предположение, является предположение о выполнимости условия Липшица
\[
\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in D,\; 0<L<\infty.
\]
%Отметим, что значение константы $L$ обычно является априори неизвестным, что делает ее оценку одной из ключевых проблем при построении методов липшицевой оптимизации.$\overline{\underline{\text{XX}}}$
Предположение липшицевости целевой функции типично для многих подходов к разработке оптимизационных алгоритмов \cite{Jones1993,Pinter1996,Zilinskas2008,Evtushenko2009}.
При этом многие известные методы основаны на различных способах разбиения области поиска на систему подобластей и последующего выбора наиболее перспективной подобласти для размещения очередного испытания (вычисления значения целевой функции) \cite{Jones2009,Zilinskas2010,Evtushenko2013,Kvasov2013,Paulavicius2016}.

Важным свойством задач глобальной оптимизации является тот факт, что в отличие от задач на поиск локального экстремума, глобальный минимум является интегральной характеристикой решаемой задачи. Чтобы убедиться, что точка $y^*\in D$ является решением задачи, недостаточно исследовать лишь ее окрестность, требуется исследование всей области поиска. Как результат, при минимизации существенно многоэкстремальных функций численный метод должен построить покрытие области поиска и число узлов этого покрытия увеличивается экспоненциально с ростом размерности. 
Эта особенность обуславливает высокую трудоемкость решения задач многоэкстремальной оптимимизации, и размерность является решающим фактором, влияющим на сложность их решения.

Для преодоления сложностей, вызванных размерностью решаемой задачи, в многоэкстремальной оптимизации широко используются подходы к уменьшению размерности.  
Например, используемое в указанных подходах симплексное или диагональное разбиение области поиска позволяет использовать для решения исходной многомерной задачи методы решения одномерных задач (see, for example, \cite{PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}). 
Известным подходом к редукции размерости является также использование Peano space-filling curves mapping the multidimensional domain onto one-dimensional interval \cite{Strongin2000}.

В данной работе мы будем использовать еще один способ, основанный на the nested optimization scheme \cite{Shi2000,Grishagin2001,VanDam2010,Grishagin2015} and in its generalization \cite{Grishagin2016,Grishagin2016_1}.
The nested optimization scheme, с одной стороны, не ухудшает свойств минимизируемой функции (в отличие от редукции с использованием Peano curves), а с другой стороны, не требует использования сложных структур данных для поддержки симплексных или диагональных разбиений допустимой области. 
Одновременно с этим схема вложненной оптимизации позволяет свести решение многомерной задачи к решению набора стандартных одномерных задач и использовать широкий спектр алгоритмов одномерной глобальной оптмизации для их решения \cite{}.

\section{Methods}

\subsection{Core global search algorithm}\label{CoreGSA}

В качестве стандартной задачи рассмотрим одномерную задачу многоэкстремальной оптимизации 
\begin{equation}\label{uni_problem}
\varphi^\ast = \varphi(x^\ast)=\min{\left\{\varphi(x):x\in \left[0,1\right] 
\right\}}
\end{equation}
с целевой функцией, удовлетворяющей условию Липшица.

Приведем описание алгоритма глобального поиска (АГП) для решения базовой задачи в соответствии с \cite{Strongin2000}.
В процессе своей работы АГП порождает последовательность точек $x^i$, в которых вычисляются значения целевой функции $z^i=\varphi(x^i)$. 
Будем называть процесс вычисления значения целевой функции \textit{испытанием}.

В соответствии с алгоритмом первые два испытания проводятся в граничных точках отрезка $[0,1]$, т.е. $x^0=0,\;x^1=1$. 
В этих точках вычисляются значения целевой функции $z^0=\varphi(x^0),\;z^1=\varphi(x^1)$ и устанавливается значение счетчика $k=1$. 
Точка очередного испытания $x^{k+1}, k\geq 1,$ выбирается в соответствии со следующими действиями.

 Step 1. Перенумеровать нижним индексом (начиная с 0) точки $x^i,\:0\leq i\leq k$, проведенных испытаний в порядке возрастания координаты, т.е.
\[
0=x_0<x_1<\ldots <x_{k}=1.
\] 
Сопоставить точкам $x_i, \; 0\leq i\leq k$, вычисленные в них значения целевой функции $z_i=\varphi(x_i), \; 0\leq i\leq k$.

Step 2. Вычислить максимальное абсолютное значение относительной первой разности
\begin{equation}\label{mu}
\mu=\max_{1\leq i\leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
\end{equation}
где $\Delta_i = x_i-x_{i-1}$. Если вычисленное в соответствии с данной формулой значение равно нулю, то положить $\mu = 1$.

Step 3. Для всех интервалов $(x_{i-1},x_i),1\leq i\leq k$,  вычислить значение
\begin{equation}\label{R}
R(i)=r\mu\Delta_i+\frac{(z_i-z_{i-1})^2}{r\mu\Delta_i}-2(z_i+z_{i-1}),
\end{equation} 
называемое \textit{характеристикой} интервала; величина $r>1$ является параметром алгоритма.

Step 4. Найти интервал $(x_{t-1},x_t)$ с максимальной характеристикой
\begin{equation}\label{MaxR}
R(t)=\max_{1\leq i\leq {k}}R(i).
\end{equation}
Если максимальная характеристика соответствует нескольким интервалам, то в качестве $t$ выбрать минимальное число, удовлетворяющее (\ref{MaxR}).

Step 5. Провести новое испытание в точке
\begin{equation}\label{xk1}
x^{k+1}=\frac{1}{2}(x_{t-1}+x_t) - \frac{z_t-z_{t-1}}{2r\mu}.
\end{equation}

Алгоритм прекращает свою работу при выполнении условия $\Delta_t<\epsilon$; здесь $t$ из (\ref{MaxR}), а $\epsilon>0$ есть заданная точность. 
В качестве оценки решения задачи выбираются значения 
\[
z_k^\ast=\min_{0\leq i \leq k}\varphi(x^i), \ x_k^\ast=\arg \min_{0\leq i \leq
 k}\varphi(x^i).
\]

Теоретические условия, определяющие сходимость алгоритма, представлены в \cite{Strongin2000}.

%Пример работы алгоритма - функция Хилла. Эту же задачу надо будет показать и в след.разделе

\subsection{Machine Learning Regression как инструмент для выделения областей притяжения локальных экстремумов}

%In this section, in order to have a self-contained description of our approach, we briefly present the Decision Tree Regression used in our experiments.

\subsection{Адаптивная схема редукции размерности }

Рекурсивная схема вложенной оптимизации основана на известном соотношении \cite{Grishagin2001} 
\begin{equation}\label{nested}
\min_{y \in D}\varphi(y) = \min_{y_1\in\left[a_1,b_1\right]}\min_{y_2\in\left[a_2,b_2\right]}...\min_{y_N\in\left[a_N,b_N\right]}\varphi(y),
\end{equation}
которое позволяет свести решение исходной многомерной задачи (\ref{main_problem}) к решению семейства рекурсивно связанных одномерных подзадач.

Для формального описания многошаговой схемы введем семейство функций, определяемых в соответствии с соотношениями 
\begin{equation}\label{nested_N}
\varphi_N(y_1,...,y_N) \equiv \varphi(y_1,...,y_N),
\end{equation}
\begin{equation}\label{nested_i}
\varphi_i(y_1,...,y_i) = \min_{ y_{i+1} \in\left[a_{i+1},b_{i+1}\right]} \varphi_{i+1}(y_1,...,y_i,y_{i+1}), 1\leq i\leq N-1.
\end{equation}

Тогда, в соответствии с (\ref{nested}), для решения многомерной задачи (\ref{main_problem}) достаточно решить одномерную задачу  
\begin{equation}\label{nested_1}
\varphi^* = \min_{y_1\in\left[a_1,b_1\right]}\varphi_1(y_1).
\end{equation}
Однако каждое вычисление значения функции $\varphi_1$ в некоторой фиксированной точке $y_1$ предполагает решение одномерной задачи оптимизации второго уровня 
\begin{equation}
\varphi_1(y_1) = \min_{y_2\in\left[a_2,b_2\right]}\varphi_2(y_1,y_2).
\end{equation}
Вычисление значений функции $\varphi_2$ в свою очередь требует одномерной минимизации функции $\varphi_3$ и т.д. вплоть до решения задачи
\begin{equation}
\varphi_{N-1}(y_1,...,y_{N-1}) = \min_{ y_{N} \in\left[a_{N},b_{N}\right]} \varphi_{N}(y_1,...,y_{N})
\end{equation}
на последнем уровне рекурсии.

Решение возникающего в схеме вложенной оптимизации множества подзадач (\ref{nested_i}) может быть организовано различными способами. 
Очевидный способ (детально проработанный в \cite{Grishagin2001,Grishagin2015} основан на решении подзадач в соответствии с рекурсивным порядком их порождения. Однако здесь возникает потеря значительной части информации о целевой функции.

Иным  подходом является адаптивная схема, в которой все подзадачи решаются одновременно, что позволяет более полно учитывать информацию о многомерной задаче и за счет этого ускорять процесс ее решения.
Данный подход был теоретически обоснован и апробирован в \cite{Grishagin2016,Grishagin2016_1,Grishagin2018}.

Кратко отметим, что в рамках исходной схемы вложенной оптимизации порождаемые подзадачи решаются строго последовательно; получаемая в результате иерархическая схема порождения и решения подзадач имеет вид дерева. Построение этого дерева происходит динамически в процессе решения исходной задачи (\ref{main_problem}). При этом вычисление одного значения функции $\varphi_i(y_1,y_2,...,y_i)$ на $i$-м уровне требует полного решения всех задач одного из поддеревьев уровня $i+1$.

Адаптивная многошаговая схема редукции размерности изменяет порядок решения подзадач: они будут решаться не по одной (в соответствии с их иерархией в дереве задач), а одновременно, т.е. будет существовать некоторое множество подзадач, находящихся в процессе решения. В рамках адаптивной схемы:
\begin{itemize}
	\item 
для вычисления значения функции $i$-го уровня из (\ref{nested_i}) порождается новая задача уровня $i+1$, в которой проводится только одно испытание, после чего новая порожденная задача включается в множество уже имеющихся задач, подлежащих решению;
	\item 
	итерация глобального поиска состоит в выборе одной (наиболее перспективной) задачи из множества имеющихся задач, в которой проводится одно испытание; точка проведения нового испытания определяется в соответствии с базовым алгоритмом глобального поиска из subsection \ref{CoreGSA};
	\item
в качестве минимальных значений функций из (\ref{nested_i}) используются их текущие оценки, полученные на основе накопленной поисковой информации.
\end{itemize}

Краткое описание основных шагов адаптивной схемы редукции размерности состоит в следующем.

Пусть вложенные подзадачи вида (\ref{nested_i}) решаются с помощью алгоритма глобального поиска, описанного в subsection \ref{CoreGSA}. Тогда каждой подзадаче (\ref{nested_i}) можно присвоить некоторое числовое значение, называемое характеристикой этой задачи. В качестве такой характеристики можно взять значение $R(t)$ из (\ref{MaxR}), т.е. максимальную характеристику из характеристик интервалов, сформированных в данной задаче. Чем выше значение данной характеристики, тем более перспективной является подзадача для продолжения поиска в ней глобального минимума исходной задачи (\ref{main_problem}). Поэтому на каждой итерации выбирается подзадача с максимальной характеристикой для проведения в ней очередного испытания. Это испытание либо приводит к вычислению значения целевой функции $\varphi(y)$ (если выбранная подзадача принадлежала уровню $j=M$), либо порождает новые подзадачи согласно (\ref{nested_i}) при $j\leq M-1$. В последнем случае новые порожденные задачи добавляются к текущему множеству задач, вычисляются их характеристики и процесс повторяется. Завершение процесса оптимизации происходит, когда в корневой задаче выполняется условие остановки алгоритма, решающего эту задачу.

\section{Experimental Results}

\section{Conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
\begin{quote}
This is an example of a quote.
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\subsection{Subsection}
\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}
\item	First bullet;
\item	Second bullet;
\item	Third bullet.
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}
\item	First item; 
\item	Second item;
\item	Third item.
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, etc.

%\begin{figure}[H]
%\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
%\caption{This is a figure. Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.\label{fig1}}
%\end{figure}

% The MDPI table float is called specialtable
\begin{specialtable}[H] 
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
%%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data\\
\bottomrule
\end{tabular}
\end{specialtable}

%\begin{listing}[H]
%\caption{Title of the listing}
%\rule{\columnwidth}{1pt}
%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%\rule{\columnwidth}{1pt}
%\end{listing}

Text.

Text.

\subsection{Formatting of Mathematical Components}

This is the example 1 of equation:
\begin{equation}
a = 1,
\end{equation}
the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

This is the example 2 of equation:
\end{paracol}
\nointerlineskip
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
\end{equation}

% Example of a figure that spans the whole page width (the commands \widefigure and \begin{paracol}{2}, \linenumbers, and\switchcolumn need to be present). The same concept works for tables, too.
%\begin{figure}[H]	
%\widefigure
%\includegraphics[width=15 cm]{Definitions/logo-mdpi}
%\caption{This is a wide figure.\label{fig2}}
%\end{figure}  
\begin{paracol}{2}
\linenumbers
\switchcolumn

Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Patents}

%This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work~reported.}

\funding{This research was funded by the Ministry of Science and Higher Education of the Russian Federation, agreement number 075-15-2020-808.}

\institutionalreview{Not applicable.}

\informedconsent{Not applicable.}

%\dataavailability{In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study. Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}. You might choose to exclude this statement if the study did not report any data.}

%\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections. This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}

\conflictsofinterest{The authors declare no conflict of interest.}

%% Optional
%\sampleavailability{Samples of the compounds ... are available from the authors.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only for journal Encyclopedia
%\entrylink{The Link to this entry published on the encyclopedia platform.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
%\abbreviations{Abbreviations}{
%The following abbreviations are used in this manuscript:\\
%
%\noindent 
%\begin{tabular}{@{}ll}
%MDPI & Multidisciplinary Digital Publishing Institute\\
%DOAJ & Directory of open access journals\\
%TLA & Three letter acronym\\
%LD & Linear dichroism
%\end{tabular}}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Optional
%\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
%\appendixstart
%\appendix
%\section{}
%\subsection{}
%The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data. Mathematical proofs of results not central to the paper can be added as an appendix.
%
%\begin{specialtable}[H] 
%%\tablesize{\scriptsize}
%\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab2}}
%%\tablesize{} % You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
%\begin{tabular}{ccc}
%\toprule
%\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
%\midrule
%Entry 1		& Data			& Data\\
%Entry 2		& Data			& Data\\
%\bottomrule
%\end{tabular}
%\end{specialtable}
%
%\section{}
%All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{paracol}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
\externalbibliography{yes}
\bibliography{bibliography}

%=====================================
% References, variant B: internal bibliography
%=====================================
%\begin{thebibliography}{999}
%% Reference 1
%\bibitem[Author1(year)]{ref-journal}
%Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
%% Reference 2
%\bibitem[Author2(year)]{ref-book1}
%Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
%% Reference 3
%\bibitem[Author3(year)]{ref-book2}
%Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
%% Reference 4
%\bibitem[Author4(year)]{ref-unpublish}
%Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} stage of publication (under review; accepted; in~press).
%% Reference 5
%\bibitem[Author5(year)]{ref-communication}
%Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
%% Reference 6
%\bibitem[Author6(year)]{ref-proceeding}
%Author 1, A.B.; Author 2, C.D.; Author 3, E.F. Title of Presentation. In Title of the Collected Work (if available), Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference; Editor 1, Editor 2, Eds. (if available); Publisher: City, Country, Year (if available); Abstract Number (optional), Pagination (optional).
%% Reference 7
%\bibitem[Author7(year)]{ref-thesis}
%Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
%% Reference 8
%\bibitem[Author8(year)]{ref-url}
%Title of Site. Available online: URL (accessed on Day Month Year).
%\end{thebibliography}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2021-06-23 2021/Entropy/entropy.tex
some phrases about regression

@@
 %In this section, in order to have a self-contained description of our approach, we briefly present the Decision Tree Regression used in our experiments.
+
+%Регрессия - известный способ построения приближения функций.
+
+Функции, рассматриваемые в рамках данного исследования, принадлежат классу липшицевых функций.
+Поэтому классические методы регрессии (например, полиномиальная регрессия, когда функция приближается полиномом заданной степени) не будут хорошо соответствовать поведению функции.
+
+Более мощным инструментом являются регрессионные сплайны.
+При построении регрессионного сплайна область определения разбивается на $K$ непересекающихся подобластей, в каждой из которых функция аппроксимируется полиномом.
+При разбиении интервала на достаточное число подобластей позволяет очень точно аппроксимировать исходную функцию.
+Однако при этом модель (сплайн) сам становится сложным для того анализа, который требуется выполнить в решаемой проблеме (выделение областей притяжения локальных экстремумов).
+
+%Здесь - почему мы не выбрали нейросети
+
+Поэтому в рамках проведенного исследования для анализа локального поведения функции мы выбрали регрессионную модель, основанную на деревьях решений.
+Построение регрессии с помощью дерева решений включает два основных этапа
+1. разбиение области поиска на $K$
 

### 2021-06-25 2021/Entropy/entropy.tex
Extended description of decision trees

@@
 
+%Алгоритмы на основе метамоделей
+
 Ряд алгоритмов основан на идее мультистарта: запуск локального поиска либо начиная с разных стартовых точек, либо с использованием разных параметров.
@@
 
-Следует отметить, что алгоритмы указанных двух классов не обеспечивают гарантированную сходимость к решению задачи и проигрывают детерминированным алгоритмам по качеству работы \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a certain set).
+Следует отметить, что алгоритмы последних двух классов не обеспечивают гарантированную сходимость к решению задачи и проигрывают детерминированным алгоритмам по качеству работы \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a certain set).
 Поэтому перспективным является использование именно детерминированных методов.
@@
 
-Для преодоления сложностей, вызванных размерностью решаемой задачи, в многоэкстремальной оптимизации широко используются подходы к уменьшению размерности.  
-Например, используемое в указанных подходах симплексное или диагональное разбиение области поиска позволяет использовать для решения исходной многомерной задачи методы решения одномерных задач (see, for example, \cite{PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
+Для преодоления сложностей, вызванных размерностью решаемой задачи, в многоэкстремальной оптимизации широко используются различные подходы к уменьшению размерности.  
+Например, симплексное или диагональное разбиение области поиска позволяет использовать для решения исходной многомерной задачи методы решения одномерных задач (see, for example, \cite{PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
 Известным подходом к редукции размерости является также использование Peano space-filling curves mapping the multidimensional domain onto one-dimensional interval \cite{Strongin2000}.
@@
 The nested optimization scheme, с одной стороны, не ухудшает свойств минимизируемой функции (в отличие от редукции с использованием Peano curves), а с другой стороны, не требует использования сложных структур данных для поддержки симплексных или диагональных разбиений допустимой области.
-Одновременно с этим схема вложненной оптимизации позволяет свести решение многомерной задачи к решению набора стандартных одномерных задач и использовать широкий спектр алгоритмов одномерной глобальной оптмизации для их решения \cite{}.
+Одновременно с этим схема вложенной оптимизации позволяет свести решение многомерной задачи к решению набора стандартных одномерных задач и использовать широкий спектр алгоритмов одномерной глобальной оптимизации для их решения \cite{}.
 
@@
 
-%In this section, in order to have a self-contained description of our approach, we briefly present the Decision Tree Regression used in our experiments.
-
-%Регрессия - известный способ построения приближения функций.
-
 Функции, рассматриваемые в рамках данного исследования, принадлежат классу липшицевых функций.
@@
 При разбиении интервала на достаточное число подобластей позволяет очень точно аппроксимировать исходную функцию.
-Однако при этом модель (сплайн) сам становится сложным для того анализа, который требуется выполнить в решаемой проблеме (выделение областей притяжения локальных экстремумов).
 
-%Здесь - почему мы не выбрали нейросети
+Построить регрессию можно также и с использованием такого мощного инструмента, как искусственные нейронные сети. 
+Для построения регрессии могут быть использованы сети разного типа, например, многослойный персептрон, сеть радиальных базисных функций, и т.д. 
+Однако в обоих этих случаях модель (сплайн или же нейронная сеть) сама становится сложной для того анализа, который требуется выполнить в решаемой проблеме (выделение областей притяжения локальных экстремумов).
 
 Поэтому в рамках проведенного исследования для анализа локального поведения функции мы выбрали регрессионную модель, основанную на деревьях решений. 
-Построение регрессии с помощью дерева решений включает два основных этапа
-1. разбиение области поиска на $K$
+Например, если целевая функция хорошо аппроксимируется полиномами, то полиномиальная регрессия, конечно же, будет адекватно передавать свойства функции.
+Однако если между имеет место более сложная зависимость, тогда дерево решений может превзойти по качеству аппроксимации классические варианты регрессии.
+%Здесь нужен рисунок - функция Шекеля, и два варианта регрессии - полином и деревья решений. Шекель должен плохо полиномами представляться, т.к. там функции вида 1/x
+Одновременно с этим регрессия на основе дерева решений позволяет с достаточной точностью легко выделять области притяжения локальных экстремумов.
 
-\subsection{Адаптивная схема редукции размерности }
+Построение регрессии с помощью дерева решений включает два основных этапа:
+\begin{itemize}
+	\item Область поиска $D$ разбивается на $J$ непересекающихся подобластей $D_1, D_2, ..., D_J$, причем $D = \bigcup_{j=1}^{J}{D_j}$.
+	\item Любому значению, попадающему в область $D_j$, т.е. $x \in D_j$, ставится в соответствие среднее значение $c_j$ по значениям training trials, попавших в эту область.
+\end{itemize}
+
+Фактически, деревья решений представляют модель функции вида
+\begin{equation}\label{dtree}
+f(x) = c_j, \; x\in D_j.
+\end{equation}
+
+Указанная модель является, с одной стороны, довольно простой, а с другой стороны, адекватно отражает интересующие нас свойства функции (наличие или отсутствие локальных минимумов).
+Выделение областей притяжения локальных минимумов по модели (\ref{dtree}) можно организовать следующим образом.
+
+%Здесь - описание алгоритма выделения локальных минимумов, с примером на Шекеле.
+
+\subsection{Адаптивная схема редукции размерности}
 

### 2021-06-25 2021/ICNAAM/2021_ICNAAM.tex

@@
 \section{INTRODUCTION}
+
+\section{GLOBAL SEARCH ALGORITHM}
+
+$y=(y_1,y_2,...,y_N)$
+
+\begin{eqnarray}\label{problem}
+&\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
+&D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.\label{D}
+\end{eqnarray}
+
+$y^*$,
+
+$y^i\in D, \; 1 \leq i \leq k$.
+
+$\varphi(y)$
+
+\[
+\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D, 0<L<\infty.
+\]
+
+\[
+\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}},
+\]
+
+$x^1,...,x^k$
+
+\begin{equation}\label{x_i}
+0=x_0<x_1<\dots <x_k<x_{k+1}=1,
+\end{equation}
+
+$z_i=\varphi(y(x_i)), \; 0 \leq i\leq k+1,$
+
+$(x_{i-1},x_i), \; 1\leq i \leq k, $
+
+\[
+R(t)=\max{R(i):1 \leq i \leq k}.
+\]
+
+$(x_{t-1},x_t)$, i.e. $x^{k+1} \in (x_{t-1},x_t)$.
+
+  $\left(x_t-x_{t-1}\right)^{1/N}<\epsilon$
+
+\section{NUMERICAL EXPERIMENTS}
+
+\[
+\nabla \cdot \vec{U} = 0
+\]
+
+\[
+\frac{\partial \vec{U}}{\partial t} + \nabla \cdot \left( \vec{U} \otimes \vec{U}\right) - \nabla \cdot v \left( \nabla \vec{U} + (\nabla \vec{U})^T\right) = - \frac{1}{\rho} \nabla p.
+\]
+
+\[
+\vec{U}(x,y) = 0, \; p(x,y) = 0.
+\]
+
+\[
+\nabla p \cdot \vec{n} = 0, \; \vec{U} = 0.
+\]
+
+\[
+\nabla p \cdot \vec{n} = 0, \; \vec{U} = \vec{U_i}.
+\]
+
+\[
+ p = 0, \; \vec{U} = \vec{U_o}.
+\]
+
+\begin{figure}%[ht]
+\includegraphics[width=1.0\linewidth]{fig1.png}
+\caption{Figure 1}
+\label{fig}
+\end{figure}
+
 Do not abbreviate Figure, Equation, etc.; display items are always singular, i.e., Figure 1 and 2.
@@
 Please do not use footnotes in the text, additional information can be added to the reference list.
-
-This is the paragraph spacing that occurs when you use the [ENTER] key.
-
-\section{FIRST-ORDER HEADING}
-This is an example of dummy text. This is an example of dummy text. This is an example of dummy text.
-This is an example of dummy text. This is an example of dummy text. This is an example of dummy text.
-
-This is the paragraph spacing that occurs when you use the [ENTER] key.
-
-% enumerate
-\noindent Numbered lists may also be included and should look like this:
-
-\begin{enumerate}
-\item This is an example of numbered listing.
-\item This is an example of numbered listing.
-\end{enumerate}
-
-\noindent Bulleted lists may be included and should look like this:
-% itemize
-
-\begin{itemize}
-\item This is an example of bulleted listing.
-\item This is an example of bulleted listing. This is an example of bulleted listing.
-\end{itemize}
-
-\noindent This is an example of single line numbered equation.
-\begin{equation}
-\frac{d[F_1]}{d\omega_2} = SAm_2\cos\omega,\frac{d[F_1]}{d\omega_3}= SAm_2\cos\omega.
-\end{equation}
-
-This is an example of a multiline numbered equation.
-% Numbered multiple equation
-\begin{eqnarray}
-p_{t_{10,1}}&=&\left(\frac{N_{cu}^2}{ N_c ^2}\right)\left(\frac{N_{ar}^2}{N_a^2}\right)\left(\frac{N_{ar}-1}{N_{ar}}\right),\\
-p_{t_{10,2}}&=&\left(\frac{N_{cu}^2}{ N_c ^2}\right)\left(\frac{N_{ar}}{N_a^2}\right).
-\end{eqnarray}
-% Unnumbered equation
-%% \begin{eqnarray*}
-%% {\rm ((Multi\ Equation\ Unnumbered))}\\
-%% {\rm ((Multi\ Equation\ Unnumbered))}\\
-%% {\rm ((Multi\ Equation\ Unnumbered))}
-%% \end{eqnarray*}
-%% this is an example of dummy text.
-% Unnumbered equation
-%% \[
-%% {\rm ((Single\ Equation\ Unnumbered))}
-%% \]
-For more on equations, please refer to the guide.
-
-% Figure
-%\begin{figure}[h]
-  %\centerline{\includegraphics[width=150pt]{fig_1.pdf}}
-  %\caption{To format a figure caption use the \LaTeX template style: Figure Caption.
-The text ``FIGURE 1,'' which labels the caption, should be bold and in upper case.
-If figures have more than one part, each part should be labeled (a), (b), etc.
-Using a table, as in the above example, helps you control the layout.}
-%\end{figure}
 
@@
 Type the word ``FIGURE 1.'' in bold uppercase, followed by a period.\footnote{This is an example of a footnote.}
-
-\subsection{Tables (Second Level Heading)}
-Due to the wide range and complexity of tables, we simply offer an example for guidance. Please follow the style for Table \ref{tab:a} (and figure) captions.
-
-% Table
-\begin{table}[h]
-\caption{To format a table caption, use the \LaTeX template style: Table Caption.
-The text ``{\bf TABLE 1,}'' which labels the caption, should be bold and all letters capitalized.
-Center this text above the table.
-Tables should have top and bottom rules, and a rule separating the column heads from the rest of the table only.}
-\label{tab:a}
-\tabcolsep7pt\begin{tabular}{lcccc}
-\hline
-  & \tch{1}{c}{b}{Single\\ outlet}  & \tch{1}{c}{b}{Small\\ multiple}  & \tch{1}{c}{b}{Large\\ multiple}  & \tch{1}{c}{b}{Total}   \\
-\hline
-1982 & 98 & 129 & 620    & 847\\
-1987 & 138 & 176 & 1000  & 1314\\
-1991 & 173 & 248 & 1230  & 1651\\
-1998 & 200 & 300 & 1500  & 2000\\
-\hline
-\end{tabular}
-\end{table}
-
-\subsection{Font Embedding (Second Level Heading)}
-As the author and creator of your article PDF, you have the most intimate knowledge of exactly what the PDF should display.
-We ask all authors to carefully check their article PDF prior to submission.
-Perform visual inspections to detect subtle font errors and ensure that all fonts are embedded.
-With the wide range of tools and software that authors use to create PDFs, and the number of devices and platforms that readers use to view/print them, font embedding by authors is not only ``nice-to-have'', it is essential.
-
-\subsubsection{Why Should I Care About Font Embedding? (Third Level Heading)}
-Embedding fonts into your PDF file is critically important for two reasons:
-\begin{itemize}
-\item Commercial printing companies are unable to print PDFs without the correct fonts embedded.
-\item To ensure that your online article PDF file displays and prints correctly for everyone who wants to read your work.
-\end{itemize}
-
-Readers of scientific articles use an ever-increasing range of devices and applications to access, view, and print PDFs  from smart phones and tablets to desktop computers running any one of a number of operating systems.
-To ensure that readers of  your article can display and print it correctly, it is important for your article's PDF file to be truly portable.
-Your PDF file needs to be fully ``self-contained.''%\vfill\pagebreak
-
-\section{FINAL KEY POINTS TO CONSIDER (FIRST LEVEL HEADING)}
-Here are the main points you need to follow (the AIP Publishing author template packages contain comprehensive guidance):
-
-\begin{itemize}
-\item Write and prepare your article using the AIP Publishing template.
-\item Create a PDF file of your paper (making sure to embed all fonts).
-\item Send the following items to your conference organizer:
-\begin{itemize}
-\item PDF file of your paper
-\item Signed Copyright Transfer Agreement
-\item (If it applies) Copies of any permissions to re-use copyrighted materials in your article (e.g., figures from books/journals)
-\end{itemize}
-\end{itemize}
-
-% Sections that will go in second font
 

### 2021-06-29 2021/Entropy/entropy.tex
Добавлена информация про выделения области притяжения локального минимума

@@
 
-Фактически, деревья решений представляют модель функции вида
+Фактически, деревья решений строят модель функции вида
 \begin{equation}\label{dtree}
@@
 Указанная модель является, с одной стороны, довольно простой, а с другой стороны, адекватно отражает интересующие нас свойства функции (наличие или отсутствие локальных минимумов).
-Выделение областей притяжения локальных минимумов по модели (\ref{dtree}) можно организовать следующим образом.
-
+Выделение областей притяжения локальных минимумов с использованием модели (\ref{dtree}) можно организовать следующим образом.
 %Здесь - описание алгоритма выделения локальных минимумов, с примером на Шекеле.
 
-\subsection{Адаптивная схема редукции размерности}
+Пусть мы имеем $x^{k+1}$ - текущая точка испытания, для данной точки необходимо найти номер подобласти  $j$, такой что $x^{k+1} \in D_j$ 
+Далее, необходимо проверить значения $c$ соответствующие соседним подобластям. Если выполняется условие 
+\begin{equation}\label{LocalPointC}
+c_{j-t_1} > … > c_{j-1} > c_{j} < c_{j+1} < c_{j+t_2}
+\end{equation}
+\begin{equation}\label{LocalPointT}
+ t_1 + t_2 + 1  = T
+\end{equation}
+, где T - параметр метода, экспериментально было выбрано значение 5.
+Тогда область $D_j$ является областью притяжения локального минимума.
+
+Для того что бы внедрить данные подходы в АГП, необходимо добавить следующие шаги в алгоритм описанные в "3.1. Core global search algorithm"
+
+Step 6.
+Построить дерево решения на основе точек проведенных испытаний, $x_0<x_1<\ldots <x_{k}$, и получить соответствующую кусочно линейную аппроксимацию: каждой подобласти $D_1, D_2, ..., D_J$ соответствует значение $с_1, с_2, ..., с_J$.
+
+Step 7.
+Для текущей точки испытания $x^{k+1}$  найти $j$, такое что $x^{k+1} \in D_j$ и проверить выполняется ли условие (\ref{LocalPointC}).
+Если условие выполняется запустить локальный метод из точки $x^{k+1}$.
+
+\subsection{Адаптивная схема редукции размерности }
 

### 2021-06-30 2021/Entropy/entropy.tex
minor corrections in Russian

--- 2021-06-25 2021/Entropy/entropy.tex
@@
 
-\subsection{Machine Learning Regression как инструмент для выделения областей притяжения локальных экстремумов}
+\subsection{Machine Learning Regression как инструмент для выделения областей притяжения локальных экстремумов}\label{TreeGSA}
 
@@
 Указанная модель является, с одной стороны, довольно простой, а с другой стороны, адекватно отражает интересующие нас свойства функции (наличие или отсутствие локальных минимумов).
-Выделение областей притяжения локальных минимумов по модели (\ref{dtree}) можно организовать следующим образом.
+Выделение областей притяжения локальных минимумов с использованием модели (\ref{dtree}) можно организовать следующим образом.
+%Здесь - описание алгоритма выделения локальных минимумов, с иллюстрацией на функции Шекеля.
 
-%Здесь - описание алгоритма выделения локальных минимумов, с примером на Шекеле.
+Пусть $x^{k+1}$ -- текущая точка испытания. Для данной точки ищется номер $j$, такой что $x^{k+1} \in D_j$ 
+Далее сравниваются значения $c_j$, соответствующие соседним подобластям. Если выполняется одно из условий
+\begin{gather}\label{LocalPointC}
+c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3} \leq c_{j+4}, \; j=1, \\ \nonumber
+c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3}, \; j=2,  \\ \nonumber
+c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2}, \; 3 \leq j \leq J-2;\\
+c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1}, \; j=J-1, \\ \nonumber
+c_{j-4} \geq c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j}, \; j=J, \nonumber
+\end{gather}
+то область $D_j$ считается областью притяжения локального минимума. В ней можно запускать локальный поиск.
 
-\subsection{Адаптивная схема редукции размерности}
+%\begin{equation}\label{LocalPointC}
+%c_{j-t_1} > ... > c_{j-1} > c_{j} < c_{j+1} < c_{j+t_2}
+%\end{equation}
+%\begin{equation}\label{LocalPointT}
+ %t_1 + t_2 + 1  = T,
+%\end{equation}
+%где $T$ -- параметр метода, экспериментально было выбрано значение 5.
+%Тогда область $D_j$ является областью притяжения локального минимума.
+
+Для того, чтобы использовать выделение локальных минимумов в алгоритме глобального поиска, описанном в subsection \ref{CoreGSA}, его необходимо дополнить следующими шагами.
+
+Step 6.
+Построить дерево решений на основе результатов выполненных испытаний, и получить соответствующую кусочно линейную аппроксимацию, которая каждой подобласти $D_1, D_2, ..., D_J$ сопоставляет значение $с_1, с_2, ..., с_J$.
+
+Step 7.
+Для точки $x^{k+1}$ текущего испытания найти $j$ такой, что $x^{k+1} \in D_j$ и проверить выполнимость условия (\ref{LocalPointC}).
+Если условие (\ref{LocalPointC}) выполняется, запустить локальный поиск из точки $x^{k+1}$.
+
+В качестве локального алгоритма на шаге 7 может использоваться любой из локальных методов нулевого порядка, например,  golden section search or parabolic interpolation \cite{Press}.
+
+%Нужно еще описать исключение интервалов из глоабльно поиска, и также - новый критерий остановки.
+
+\subsection{Адаптивная схема редукции размерности }
 
@@
 	\item 
-	итерация глобального поиска состоит в выборе одной (наиболее перспективной) задачи из множества имеющихся задач, в которой проводится одно испытание; точка проведения нового испытания определяется в соответствии с базовым алгоритмом глобального поиска из subsection \ref{CoreGSA};
+	итерация глобального поиска состоит в выборе одной (наиболее перспективной) задачи из множества имеющихся задач, в которой проводится одно испытание; точка проведения нового испытания определяется в соответствии с базовым алгоритмом глобального поиска из subsection \ref{CoreGSA} или же модифицированным алгоритмом из subsection \ref{TreeGSA};
 	\item
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``Conceptualization, X.X.
-and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
-All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
-Authorship must be limited to those who have contributed substantially to the work~reported.}
+\authorcontributions{Conceptualization and methodology, K.B.; software and validation, I.L.; formal analysis, K.B.; investigation, I.L.; data curation, I.L.; writing---original draft preparation, K.B.; writing---review and editing, K.B.; visualization, I.L.; funding acquisition, K.B.
+All authors have read and agreed to the published version of the manuscript.}
 

### 2021-07-01 2021/ICNAAM/2021_ICNAAM.tex
Add some references from ISP RAN
first version in English

@@
 % Title portion
-\title{The Title Goes Here with Each Initial Letter Capitalized}
+\title{Modelling and Optimization of Injector Parameters with OpenFOAM and Globalizer Tools}
 
@@
 \begin{abstract}
-The AIP Proceedings article template has many predefined paragraph styles for you to use/apply as you write your paper.
-To format your abstract, use the \LaTeX template style: {\itshape Abstract.} Each paper must include an abstract.
-Begin the abstract with the word ``Abstract'' followed by a period in bold font, and then continue with a normal 9 point font.
+The paper presents the results of solving the problem of modelling and research of processes taking place in a fuel system injector.
+Modelling has been carried using the tools of the OpenFOAM system.
+The problem of differential pressure minimization at the injector inlet and outlet was set and solved.
+This problem was a Lipschitz optimization problem with a black-box type objective function.
+To solve it, we used an efficient global search algorithm implemented in the Globalizer system.
 \end{abstract}
@@
 
+The problem of designing new engineering objects is inseparably connected with the problem of selecting the best values of item parameters, because the choice of specific values of parameters (with a fixed overall structure) can significantly affect its characteristics.
+
+As engineering objects  and conditions of their functioning become more complicated, it results in a significant increase in the complexity of corresponding mathematical models and numerical methods for their analysis.
+Nowadays, the main (and often the only possible) tool for such analysis is supercomputer simulation of the object's behaviour.
+For this purpose, CAD/CAM/CAE-systems are widely used.
+A well-known example of open source software of this kind is OpenFOAM, an open platform for numerical modelling of continuum mechanics problems \cite{OpenFOAM}).
+
+In this study, we used the OpenFOAM system to simulate a fuel injector.
+In the simulation process, the pressure drop at the inlet and outlet of the injector was calculated.
+The problem of minimizing the pressure drop in order to save fuel by reducing the concentration of combustible substance in the mixture was set and solved.
+The parameters used in the problem included the width of the gap between the injector inlet and the injector outlet.
+
 \section{GLOBAL SEARCH ALGORITHM}
 
-$y=(y_1,y_2,...,y_N)$
-
+The global optimization problem
 \begin{eqnarray}\label{problem}
@@
 
-$y^*$,
+can serve as a mathematical model for the process of choosing optimum values of parameters $y=(y_1,y_2,...,y_N)$ of the engineering object under consideration.
+The choice of the value  $y^*$ that determines the best option  is reduced to an approximate (with some required accuracy) solution of the above optimization problem based on the evaluation of object characteristics for the set of options $y^i\in D, \; 1 \leq i \leq k$, determined by available resources.
 
-$y^i\in D, \; 1 \leq i \leq k$.
+The optimal parameter selection problems under consideration are characterized by the fact that the objective function $\varphi(y)$ is not defined analytically; there is only some algorithm to calculate its values at the points of the domain $D$.
+In this case, one search trial (calculating the function value in a particular point of the search domain) is a computationally intensive operation \cite{Kalyulin2017}.
 
-$\varphi(y)$
+Multi-extremal optimization problems are essentially more time-consuming in comparison with other types of optimization problems because global optimum is an integral characteristic of the problem to be solved and demands investigation of the whole search domain.
+As a result, the search for the global optimum is reduced to generating  some covering (grid) in the parameter domain and choosing the best function value on this grid.
+The amount of calculations can be reduced by creating a non-uniform coverage of the search domain: the grid has to be dense enough in the vicinity of the global optimum and more sparse farther away from the solution being sought.
 
+A typical assumption used in many global optimization methods \cite{Evtushenko2013,Jones2009,Zilinskas2010,Pinter1996} is the assumption that the objective function $\varphi(y)$ satisfies the Lipschitz condition
 \[
@@
 
+At the same time.
+various dimensionality reduction schemes are widely used to reduce the complexity of designing global optimization algorithms.
+By means of  such schemes, it is possible to reduce solution of multidimensional optimization problems to the solution of one or more nested one-dimensional optimization problems.
+
+One approach to dimensionality reduction is to use Peano-type space filling curves (\textit{evolvents}).
+Using evolvents $y(x)$, which continuously and uniquely map a unit interval onto a hypercube, reduces the problem of finding the minimum of the function $\varphi(y)$ in the domain $D$ to finding the minimum of a one-dimensional function  on the interval $[0,1]$
 \[
-\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}},
+\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}}.
 \]
 
-$x^1,...,x^k$
+The above assumptions and approaches have led to the development of a number of global search algorithms based on the information-statistical approach \cite{Strongin2000}.
+These algorithms are implemented in the Globalizer software system for solving global optimization problems \cite{globalizerSystem}.
+Here, we present a scheme of the basic global search algorithm (GSA).
+This is a Divide-the-Best algorithm, i.e.
+it is aimed to conduct a subsequent search trial in the most promising (from the point of view of global minimum detection) search subdomain.
 
+The first two trials are performed at the boundary points of the search domain $x^0 = 0$ and $x^1 = 1$.
+The selection of the point  $x^{k+1}, \; k \geq 1,$  of any next trial is carried out by the following steps.
+
+Step 1. Renumber the points  $x^1,...,x^k$ of the preceding trials by the lower indices in increasing order of the coordinate values, i.e.
 \begin{equation}\label{x_i}
@@
 \end{equation}
+and juxtapose to them the values $z_i=\varphi(y(x_i)), \; 0 \leq i\leq k+1,$  computed at these points.
 
-$z_i=\varphi(y(x_i)), \; 0 \leq i\leq k+1,$
+Step 2. For each interval $(x_{i-1},x_i), \; 1\leq i \leq k,$ compute the characteristics $R(i)$ using some formulae.
 
-$(x_{i-1},x_i), \; 1\leq i \leq k, $
-
+Step 3. Find the interval $(x_{t-1},x_t),$ with the maximum characteristic 
 \[
@@
 
-$(x_{t-1},x_t)$, i.e. $x^{k+1} \in (x_{t-1},x_t)$.
+Step 4. Execute the next trial in the inner point of the interval  $(x_{t-1},x_t)$, i.e. $x^{k+1} \in (x_{t-1},x_t)$.
 
-  $\left(x_t-x_{t-1}\right)^{1/N}<\epsilon$
+Step 5. Check termination condition  $\left(x_t-x_{t-1}\right)^{1/N}<\epsilon$, where $t$   is the number of the interval with the maximum characteristic and $\epsilon>0$ is the predefined accuracy.
+
+A detailed description of the algorithm, its various modifications (including parallel ones), and the corresponding convergence  theory are presented in \cite{Strongin2000,Barkalov2010,Strongin2020}.
 
@@
 
+The calculations were performed on the computer cluster of the Institute for Systems Programming and the UniHub web-lab.
+The web-lab architecture is based on the cloud computing model where resources (servers, networks, storage systems, applications, etc.) are provided remotely as a set of different-level services with on-demand servicing.
+
+To carry out numerical experiments, we used the OpenFOAM software package \cite{OpenFOAM}.
+It is a set of tools for representing, discretizing and manipulating hydrodynamic fields and solving systems of algebraic equations.
+The OpenFOAM serves to perform a complete cycle of numerical solution of continuum mechanics problems using the finite volume method: from mathematical formulation by setting initial conditions, boundary conditions, computational geometry and computational grid to representation of the results in the form of hydrodynamic fields.
+
+In our particular case, each iteration of the optimal solution search process was a separate numerical calculation of the hydrodynamic fields.
+The injector modelling was divided into several key steps.
+
+In our particular case, each iteration of the optimal solution search process was a separate numerical calculation of the hydrodynamic fields.
+The injector modelling was divided into several key steps:
+\begin{enumerate}
+\item Mathematical formulation.
+This includes a procedure for selecting equations describing the dynamics of a fluid or gas, setting boundary conditions and initial conditions.
+It does not change from iteration to iteration.
+\item Parametrization. Selection of varying parameters that will vary from iteration to iteration.
+\item Generating the geometry and the computational grid. In line with parametrization, a procedure for renewing the geometry and computational grid is performed.
+\item Calculation of hydrodynamic fields. At each iteration, the system of equations describing fluid dynamics is solved numerically.
+\item Interpretation. Physical interpretation of the results after the optimization procedure.
+\end{enumerate}
+
+The fluid dynamics for the problem described in this paper is defined by the Navier-Stokes equations for an incompressible fluid:
 \[
-\nabla \cdot \vec{U} = 0
+\nabla \cdot \vec{U} = 0,
 \]
-
 \[
-\frac{\partial \vec{U}}{\partial t} + \nabla \cdot \left( \vec{U} \otimes \vec{U}\right) - \nabla \cdot v \left( \nabla \vec{U} + (\nabla \vec{U})^T\right) = - \frac{1}{\rho} \nabla p.
+\frac{\partial \vec{U}}{\partial t} + \nabla \cdot \left( \vec{U} \otimes \vec{U}\right) - \nabla \cdot v \left( \nabla \vec{U} + (\nabla \vec{U})^T\right) = - \frac{1}{\rho} \nabla p,
 \]
-
+where $\vec{U}$ is the velocity vector, $v$  is the liquid viscosity (assumed constant), $p$  is the overpressure, $\rho$ is the density of the liquid.
+In this system of equations, there are two unknown variables: $\vec{U}$  and $p$.
+For each of these variables, initial conditions are set:
 \[
@@
 \]
-
+There are three types of boundaries in the computational domain; special boundary conditions are set on each of them. 
+On a rigid wall:
 \[
@@
 \]
-
+At the entrance to the computational domain:
 \[
@@
 \]
-
+At the exit of the computational domain:
 \[
@@
 
+The geometric representation of the computational domain is shown in Figure 1, with the parameterization points highlighted in red.
+The geometry is parametrized by increasing the distance between the red points symmetrically with respect to the channel centreline.
+
 \begin{figure}%[ht]
 \includegraphics[width=1.0\linewidth]{fig1.png}
-\caption{Figure 1}
+\caption{Schematic representation of the computational domain, with the geometry parameterization points marked in red and the black points independent of the optimization iterations.
+The grey lines show the possible deformation of the geometry.}
 \label{fig}
@@
 
-Do not abbreviate Figure, Equation, etc.; display items are always singular, i.e., Figure 1 and 2.
-Equations are always singular, i.e., Equation 1 and 2, and should be inserted using the Equation Editor, not as graphics, in the main text.
-Display items and captions should be inserted after the reference section.
-Please do not use footnotes in the text, additional information can be added to the reference list.
-
-Cite all figures in the text consecutively.
-The word ``Figure'' should be spelled out if it is the first word of the sentence and abbreviated as ``Fig.'' elsewhere in the text.
-Place the figures as close as possible to their first mention in the text at the top or bottom of the page with the figure caption positioned below the figure, all centered.
-Figures must be inserted in the text and may not follow the Reference section.
-Set figure captions in 9 point size, Times Roman font.
-Type the word ``FIGURE 1.'' in bold uppercase, followed by a period.\footnote{This is an example of a footnote.}
+The calculation was carried out with the standard pisoFoam solver and the PISO algorithm \cite{Issa1986_2}.
+A total of 300 finite elements were involved in the calculation.
+The process of finding the effective channel thickness took 3395 seconds on an Intel(R) Xeon(R) CPU X5670 2.93 GHz, during which time 69 iterations of the global search algorithm were performed.
+The horizontal component of the velocity at the entrance to the computational domain $\vec{U_i}$  was 10 m/s.
+The pressure at the exit of the computational domain $p_o$ was equal to zero, just like all components of the velocity vector at the exit $\vec{U_o}$ .
+Linear dimensions of the computational domain are shown in the schematic of Figure 1.
+As a result of solving the problem, the optimum channel thickness (0.59 cm) was found, giving the minimum fuel consumption.
 
@@
 
-\cite{Strongin2000,Kalyulin2017,Evtushenko2013,Jones2009,Zilinskas2010,Pinter1996,globalizerSystem,Barkalov2010,Strongin2020}
+\cite{Issa1986_1}
 

### 2021-07-05 2021/Entropy/entropy.tex

@@
 
-Для того, чтобы использовать выделение локальных минимумов в алгоритме глобального поиска, описанном в subsection \ref{CoreGSA}, его необходимо дополнить следующими шагами.
+Для того, чтобы использовать выделение локальных минимумов в алгоритме глобального поиска, описанном в subsection \ref{CoreGSA}, всем точкам получаемым в ходе работы алгоритма присваивается метка q, со значениями global - по умолчанию, означает что точка получена в результате описанного алгоритма глобального поиска, local - точка получена в результате работы локального метода, localMin - точка является найденным локальным минимумом.
+Индексы метки соответствуют индексам координаты $x$, т.е.
+у точки  $x^{k+1}$ метка $q^{k+1}$ и так далее.
+
+Также алгоритм необходимо дополнить следующими шагами.
 
 Step 6.
+Если хотя бы одна из точек $x_{t-1}$ или $x_t$ имеет метку local или localMin, то точка $x^{k+1}$ исключается из поисковой информации и переходим к проверке критерия остановки.
+Иначе переходим на Step 7.
+
+Step 7.
 Построить дерево решений на основе результатов выполненных испытаний, и получить соответствующую кусочно линейную аппроксимацию, которая каждой подобласти $D_1, D_2, ..., D_J$ сопоставляет значение $с_1, с_2, ..., с_J$.
 
-Step 7.
+Step 8.
 Для точки $x^{k+1}$ текущего испытания найти $j$ такой, что $x^{k+1} \in D_j$ и проверить выполнимость условия (\ref{LocalPointC}).
 Если условие (\ref{LocalPointC}) выполняется, запустить локальный поиск из точки $x^{k+1}$.
+Все $N_{local}$ точек полученных локальным методом сохраняются и получают номера $x^{k+2}, ..., x^{k+N_{local}+1}$  и используются на последующих итерациях алгоритма.
+Также эти точки получают метку local, $q^{i}=local, i = k+2, ...
+, k+N_{local}+1 $.
+Точка соответствующая найденному локальному минимуму получает метку localMin.
 
 В качестве локального алгоритма на шаге 7 может использоваться любой из локальных методов нулевого порядка, например,  golden section search or parabolic interpolation \cite{Press}.
+
+Критерий остановки будет выглядеть следующим образом:
+ Алгоритм прекращает свою работу при выполнении одного из условий
+
+1. $\Delta_t<\epsilon$; здесь $t$ из (\ref{MaxR}), а $\epsilon>0$ есть заданная точность.
+
+2. $|x^{k+1}_i-x_{t-1}|<\epsilon$ и $x_{t-1}$ имеет метку localMin.
+
+3. $|x^{k+1}_i-x_{t}|<\epsilon$ и $x_{t}$ имеет метку localMin.
 

### 2021-07-05 2021/Entropy/entropy.tex
Добавлены таблицы

@@
 
+Numerical experiments were performed on the Lobachevsky supercomputer of the University of Nizhni Novgorod (operating system -- CentOS 6.4, management system -- SLURM).
+One supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM.
+The CPU is 8-core (i.e.
+a total of 16 CPU cores are available on the node).
+
+В работе \cite{Gergel2019} описан GCGen-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: размерностью задачи, точкой глобального минимума, значением функции в ней и т.п.
+Испытания будем проводить на функциях Шекеля и Хилла-Гибсона.
+
+\begin{table}[]
+	\caption{Average number of iterations, Shekel problem}\label{table:average_Shekel}
+	\center
+\begin{tabular}{cccccc}
+\hline\noalign{\smallskip}
+                                                    & \textbf{0,0001} & \textbf{0,0005} & \textbf{0,001} & \textbf{0,005} & \textbf{0,01} \\
+DiRect                                              & 64,22                                   &                                         &                                        &                                        &                                       \\
+Базовый АГП                                         & 106                                     & 64                                      & 53                                     & 37                                     & 31                                    \\
+АГП с деревом решений & 49                                      & 45                                      & 43                                     & 38                                     & 35
+
+\end{tabular}
+\end{table}
+
+\begin{table}[]
+	\caption{Average number of iterations Hill problem}\label{table:average_Hill}
+	\center
+\begin{tabular}{cccccc}
+\hline\noalign{\smallskip}
+                      & \textbf{0,0001} & \textbf{0,0005} & \textbf{0,001} & \textbf{0,005} & \textbf{0,01} \\
+DiRect                & 65,74 (91)                              &                                         &                                        &                                        &                                       \\
+Базовый АГП           & 130                                     & 87                                      & 75                                     & 52                                     & 43                                    \\
+АГП с деревом решений & 64                                      & 60                                      & 59                                     & 55                                     & 50
+\end{tabular}
+\end{table}
+
+В таблицах \ref{table:average_Shekel} и \ref{table:average_Hill} приведено среднее число итераций полученное на  задачах Шекеля и Хилла, для алгоритмов Direct, АГП и АГП с использованием локального метода запущенного по результатам работы дерева решений.
+
+In the paper \cite{Gaviano2003} a GKLS generator is described that allows one to generate multiextremal optimization problems with previously known properties: the number of local minima, the size of their regions of attraction, the global minimum point, the value of the function in it, etc.
+
+\begin{table}[]
+	\caption{Average number of iterations GKLS problem}\label{table:average_GKLS}
+	\center
+\begin{tabular}{ccccccc}
+\hline\noalign{\smallskip}
+
+                      & \multicolumn{2}{l}{2}                                                          & \multicolumn{2}{l}{3}                                                          & \multicolumn{2}{l}{4}                                  \\
+                      & \textbf{0,002} & \textbf{0,01} & {FFFF00}\textbf{0,002} & \textbf{0,01} & \textbf{0,002} & \textbf{0,01} \\
+DiRect                &                                        & 80,04 (41)                            &                                        &                                       &                &                                       \\
+Базовый АГП           & 1820                                   & 841                                   & 53251,32                               & 8905,03                               & 665013,15      & 135972,78 (99)                        \\
+Адаптивная схема      & 1489,01                                & 937,20                                & 69764,39                               & 12716,21                              & 583903,06      & 206869,83                             \\
+АГП с деревом решений & 831,77                                 & 653,94                                & 10776,87                               & 9204,16                               & 173155,3       & 156190,43
+\end{tabular}
+\end{table}
+
+В таблицах \ref{table:average_GKLS} приведено среднее число итераций полученное на  задачах GKLS, для алгоритмов Direct, АГП, АГП с адаптивной схмемой и АГП с адаптивной схемой с использованием локального метода запущенного по результатам работы дерева решений.
+
 \section{Conclusions}

### 2021-07-05 2021/Entropy/entropy.tex
Поправлен текст с описание экспериментов

@@
 
-В таблицах \ref{table:average_Shekel} и \ref{table:average_Hill} приведено среднее число итераций полученное на  задачах Шекеля и Хилла, для алгоритмов Direct, АГП и АГП с использованием локального метода запущенного по результатам работы дерева решений.
+В таблицах \ref{table:average_Shekel} и \ref{table:average_Hill} приведено среднее число итераций полученное при решении серии из 100 задач Шекеля и Хилла, для алгоритмов Direct, АГП и АГП с использованием локального метода запущенного по результатам работы дерева решений.
+Здесь и далее в скобках указано сколько задач решилось их 100.
 
@@
 
-В таблицах \ref{table:average_GKLS} приведено среднее число итераций полученное на  задачах GKLS, для алгоритмов Direct, АГП, АГП с адаптивной схмемой и АГП с адаптивной схемой с использованием локального метода запущенного по результатам работы дерева решений.
+В таблицах \ref{table:average_GKLS} приведено среднее число итераций полученное на  задачах GKLS, для алгоритмов Direct, АГП, АГП с адаптивной схемой и АГП с адаптивной схемой с использованием локального метода запущенного по результатам работы дерева решений.
 

### 2021-07-05 2021/Entropy/entropy.tex
Поправлено оформление таблицы

@@
 
-\begin{table}[]
+\begin{specialtable}[H] 
 	\caption{Average number of iterations, Shekel problem}\label{table:average_Shekel}
@@
 \begin{tabular}{cccccc}
-\hline\noalign{\smallskip}
+\toprule
                                                     & \textbf{0,0001} & \textbf{0,0005} & \textbf{0,001} & \textbf{0,005} & \textbf{0,01} \\
+\midrule													
 DiRect                                              & 64,22                                   &                                         &                                        &                                        &                                       \\
 Базовый АГП                                         & 106                                     & 64                                      & 53                                     & 37                                     & 31                                    \\
-АГП с деревом решений & 49                                      & 45                                      & 43                                     & 38                                     & 35
+АГП с деревом решений & 49                                      & 45                                      & 43                                     & 38                                     & 35                                   \\
 
+\bottomrule
 \end{tabular}
-\end{table}
+\end{specialtable}
 
-\begin{table}[]
+\begin{specialtable}[H] 
 	\caption{Average number of iterations Hill problem}\label{table:average_Hill}
@@
 \begin{tabular}{cccccc}
-\hline\noalign{\smallskip}
+\toprule
                       & \textbf{0,0001} & \textbf{0,0005} & \textbf{0,001} & \textbf{0,005} & \textbf{0,01} \\
+\midrule					  
 DiRect                & 65,74 (91)                              &                                         &                                        &                                        &                                       \\
 Базовый АГП           & 130                                     & 87                                      & 75                                     & 52                                     & 43                                    \\
-АГП с деревом решений & 64                                      & 60                                      & 59                                     & 55                                     & 50
+АГП с деревом решений & 64                                      & 60                                      & 59                                     & 55                                     & 50                                    \\
+\bottomrule
 \end{tabular}
-\end{table}
+\end{specialtable}
 
@@
 
-\begin{table}[]
+\begin{specialtable}[H] 
 	\caption{Average number of iterations GKLS problem}\label{table:average_GKLS}
@@
 \hline\noalign{\smallskip}
+\toprule
 
-                      & \multicolumn{2}{l}{2}                                                          & \multicolumn{2}{l}{3}                                                          & \multicolumn{2}{l}{4}                                  \\
-                      & \textbf{0,002} & \textbf{0,01} & {FFFF00}\textbf{0,002} & \textbf{0,01} & \textbf{0,002} & \textbf{0,01} \\
+                      & \multicolumn{2}{l}{2D}                                                          & \multicolumn{2}{l}{3D}                                                          & \multicolumn{2}{l}{4D}                                  \\
+                      & \textbf{0,002} & \textbf{0,01} & \textbf{0,002} & \textbf{0,01} & \textbf{0,002} & \textbf{0,01} \\
+					  \midrule
 DiRect                &                                        & 80,04 (41)                            &                                        &                                       &                &                                       \\
@@
 Адаптивная схема      & 1489,01                                & 937,20                                & 69764,39                               & 12716,21                              & 583903,06      & 206869,83                             \\
-АГП с деревом решений & 831,77                                 & 653,94                                & 10776,87                               & 9204,16                               & 173155,3       & 156190,43
+АГП с деревом решений & 831,77                                 & 653,94                                & 10776,87                               & 9204,16                               & 173155,3       & 156190,43                              \\
+\bottomrule
 \end{tabular}
-\end{table}
+\end{specialtable}
 

### 2021-07-05 2021/ICNAAM/2021_ICNAAM.tex
Update references

@@
 
-The calculations were performed on the computer cluster of the Institute for Systems Programming and the UniHub web-lab.
+The calculations were performed on the computer cluster of the Institute for Systems Programming and the UniHub web-lab \cite{UniHub}.
 The web-lab architecture is based on the cloud computing model where resources (servers, networks, storage systems, applications, etc.) are provided remotely as a set of different-level services with on-demand servicing.
@@
 \[
-\vec{U}(x,y) = 0, \; p(x,y) = 0.
+\vec{U}(x,y) = 0,
+\]
+\[
+p(x,y) = 0.
 \]
@@
 \[
-\nabla p \cdot \vec{n} = 0, \; \vec{U} = 0.
+\nabla p \cdot \vec{n} = 0,
+\]
+\[
+\vec{U} = 0.
 \]
@@
 \[
-\nabla p \cdot \vec{n} = 0, \; \vec{U} = \vec{U_i}.
+\nabla p \cdot \vec{n} = 0,
+\]
+\[
+\vec{U} = \vec{U_i}.
 \]
@@
 \[
- p = 0, \; \vec{U} = \vec{U_o}.
+ p = 0,
+\]
+\[
+\vec{U} = \vec{U_o}.
 \]
@@
 
-The calculation was carried out with the standard pisoFoam solver and the PISO algorithm \cite{Issa1986_2}.
+The calculation was carried out with the standard pisoFoam solver and the PISO algorithm \cite{Issa1986_2,Issa1986_1}.
 A total of 300 finite elements were involved in the calculation.
@@
 
-\cite{Issa1986_1}
-
 % References

### 2021-07-06 2021/ICNAAM/2021_ICNAAM.tex
Update from ISP RAS

@@
 \usepackage{graphicx}
+\usepackage{tikz}
+\usetikzlibrary{shapes.geometric}
+\usetikzlibrary{positioning}
+\usetikzlibrary{patterns}
+\usetikzlibrary{arrows,decorations.markings}
+\usetikzlibrary{spy}
 
@@
 % Title portion
-\title{Modelling and Optimization of Injector Parameters with OpenFOAM and Globalizer Tools}
+\title{Simulation and Optimization of Injector Parameters with OpenFOAM and Globalizer Tools}
 
@@
 \begin{abstract}
-The paper presents the results of solving the problem of modelling and research of processes taking place in a fuel system injector.
-Modelling has been carried using the tools of the OpenFOAM system.
+The paper presents the results of solving the problem of simulation and research of processes taking place in a fuel system injector.
+Simulation has been carried using the tools of the OpenFOAM system.
 The problem of differential pressure minimization at the injector inlet and outlet was set and solved.
@@
 As engineering objects and conditions of their functioning become more complicated, it results in a significant increase in the complexity of corresponding mathematical models and numerical methods for their analysis.
-Nowadays, the main (and often the only possible) tool for such analysis is supercomputer simulation of the object's behaviour.
+Nowadays, the main (and often the only possible) tool for such analysis is supercomputer simulation of the object's behavior.
 For this purpose, CAD/CAM/CAE-systems are widely used.
-A well-known example of open source software of this kind is OpenFOAM, an open platform for numerical modelling of continuum mechanics problems \cite{OpenFOAM}).
+A well-known example of open source software of this kind is OpenFOAM, an open platform for numerical simulation of continuum mechanics problems \cite{OpenFOAM}).
 
@@
 In the simulation process, the pressure drop at the inlet and outlet of the injector was calculated.
-The problem of minimizing the pressure drop in order to save fuel by reducing the concentration of combustible substance in the mixture was set and solved.
+The problem of minimizing the pressure drop in order to save fuel by reducing the concentration of combustible substances in the mixture after leaving the ejector was set and solved.
 The parameters used in the problem included the width of the gap between the injector inlet and the injector outlet.
@@
 To carry out numerical experiments, we used the OpenFOAM software package \cite{OpenFOAM}.
-It is a set of tools for representing, discretizing and manipulating hydrodynamic fields and solving systems of algebraic equations.
-The OpenFOAM serves to perform a complete cycle of numerical solution of continuum mechanics problems using the finite volume method: from mathematical formulation by setting initial conditions, boundary conditions, computational geometry and computational grid to representation of the results in the form of hydrodynamic fields.
+It is a set of tools for representing, discretizing and manipulating hydrodynamic fields.
+OpenFOAM also provides methods for solving systems of algebraic equations.
+The OpenFOAM serves to perform a complete cycle of numerical solution of continuum mechanics problems using the finite volume method: from the mathematical formulation by setting initial conditions, boundary conditions, computational geometry, and computational grid to a representation of the results in the form of hydrodynamic fields.
 
 In our particular case, each iteration of the optimal solution search process was a separate numerical calculation of the hydrodynamic fields.
-The injector modelling was divided into several key steps.
-
-In our particular case, each iteration of the optimal solution search process was a separate numerical calculation of the hydrodynamic fields.
-The injector modelling was divided into several key steps:
+The injector simulation was divided into several key steps:
 \begin{enumerate}
@@
 It does not change from iteration to iteration.
-\item Parametrization. Selection of varying parameters that will vary from iteration to iteration.
-\item Generating the geometry and the computational grid. In line with parametrization, a procedure for renewing the geometry and computational grid is performed.
+\item Parameterization. Selection of varying parameters that will vary from iteration to iteration.
+\item Generating the geometry and the computational grid. In line with Parameterization, a procedure for renewing the geometry and computational grid is performed.
 \item Calculation of hydrodynamic fields. At each iteration, the system of equations describing fluid dynamics is solved numerically.
@@
 \]
-where $\vec{U}$ is the velocity vector, $v$  is the liquid viscosity (assumed constant), $p$  is the overpressure, $\rho$ is the density of the liquid.
+where $\vec{U}$ is the velocity vector, $\nu$  is the liquid viscosity (assumed constant), $p$  is the overpressure, $\rho$ is the density of the liquid.
 In this system of equations, there are two unknown variables: $\vec{U}$  and $p$.
@@
 \begin{figure}%[ht]
-\includegraphics[width=1.0\linewidth]{fig1.png}
+%\includegraphics[width=1.0\linewidth]{fig1.png}
+
+    \centering
+    \begin{tikzpicture}[scale=0.5,>=stealth]
+         \draw (-5,-2.5,0)--(-15,-2.5,0)--(-15,2.5,0)--(-5,2.5,0.0);%--(0,0.5,0.0)--(5,2.5,0)--(15,2.5,0)--(15,-2.5,0);
+         \draw (-5,-2.5) .. controls (0,0) .. (5,-2.5);
+         \draw (5,2.5,0)--(15,2.5,0)--(15,-2.5,0)--(5,-2.5,0.0);
+         \draw (-5,2.5) .. controls (0,0) .. (5,2.5);
+
+         \filldraw [black] (-5,-2.5,0) circle (4pt);
+         \filldraw [black] (-15,-2.5,0) circle (4pt);
+         \filldraw [black] (-15,2.5,0) circle (4pt);
+         \filldraw [black] (-5,2.5,0.0) circle (4pt);
+
+         \filldraw [black] (5,2.5,0) circle (4pt);
+         \filldraw [black] (15,2.5,0) circle (4pt);
+         \filldraw [black] (15,-2.5,0) circle (4pt);
+         \filldraw [black] (5,-2.5,0.0) circle (4pt);
+
+         \filldraw [red] (0,0.6,0.0) circle (4pt);
+         \filldraw [red] (0,-0.6,0.0) circle (4pt);
+
+         \draw [->,ultra thick] (-16,0,0) --node[above right]{$x$} (-13,0,0);
+         \draw [->,ultra thick] (-16,0,0) --node[above left]{$y$} (-16,3,0);
+
+         \draw [<->, thick] (-15,-4,0)--node[below]{$6\;\;cm$} (15,-4,0);
+
+         \draw [dashed] (-15,-2.5) -- (-15,-5);
+         \draw [dashed] (15,-2.5) -- (15,-5);
+
+         \draw [dashed] (15,-2.5) -- (16.5,-2.5);
+         \draw [dashed] (15,2.5) -- (16.5,2.5);
+
+         \draw [<->, thick] (16,-2.5,0)--node[above, rotate=-90]{$0.5\;\;cm$} (16,2.5,0);
+
+         \draw[dashed] (-17,0,0)--(17,0,0);
+
+         \draw[opacity=0.2] (-5,2.5) .. controls (0,1) .. (5,2.5);
+         \draw[opacity=0.2] (-5,-2.5) .. controls (0,-1) .. (5,-2.5);
+
+         \filldraw [red,opacity=0.2] (0,1.4,0.0) circle (4pt);
+         \filldraw [red,opacity=0.2] (0,-1.4,0.0) circle (4pt);
+
+         \draw[opacity=0.2] (-5,2.5) .. controls (0,2) .. (5,2.5);
+         \draw[opacity=0.2] (-5,-2.5) .. controls (0,-2) .. (5,-2.5);
+
+         \filldraw [red,opacity=0.2] (0,2.15,0.0) circle (4pt);
+         \filldraw [red,opacity=0.2] (0,-2.15,0.0) circle (4pt);
+
+         \draw[opacity=0.2] (-5,2.5) .. controls (0,3) .. (5,2.5);
+         \draw[opacity=0.2] (-5,-2.5) .. controls (0,-3) .. (5,-2.5);
+
+         \filldraw [red,opacity=0.2] (0,2.9,0.0) circle (4pt);
+         \filldraw [red,opacity=0.2] (0,-2.9,0.0) circle (4pt);
+
+        \draw[opacity=0.2] (-5,2.5) .. controls (0,4) .. (5,2.5);
+        \draw[opacity=0.2] (-5,-2.5) .. controls (0,-4) .. (5,-2.5);
+
+        \filldraw [red,opacity=0.2] (0,3.65,0.0) circle (4pt);
+        \filldraw [red,opacity=0.2] (0,-3.65,0.0) circle (4pt);
+
+    \end{tikzpicture}
 \caption{Schematic representation of the computational domain, with the geometry parameterization points marked in red and the black points independent of the optimization iterations.

### 2021-07-07 2021/ICNAAM/2021_ICNAAM.tex
final ICNAAM paper

@@
 &\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
-&D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.\label{D}
+&D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\},\label{D}
 \end{eqnarray}
-
 can serve as a mathematical model for the process of choosing optimum values of parameters $y=(y_1,y_2,...,y_N)$ of the engineering object under consideration.
@@
 
-At the same time.
-various dimensionality reduction schemes are widely used to reduce the complexity of designing global optimization algorithms.
+At the same time, various dimensionality reduction schemes are widely used to reduce the complexity of designing global optimization algorithms.
 By means of  such schemes, it is possible to reduce solution of multidimensional optimization problems to the solution of one or more nested one-dimensional optimization problems.
@@
 \[
-R(t)=\max{R(i):1 \leq i \leq k}.
+R(t)=\max\{R(i):1 \leq i \leq k+1\}.
 \]
@@
 \[
-\frac{\partial \vec{U}}{\partial t} + \nabla \cdot \left( \vec{U} \otimes \vec{U}\right) - \nabla \cdot v \left( \nabla \vec{U} + (\nabla \vec{U})^T\right) = - \frac{1}{\rho} \nabla p,
+\frac{\partial \vec{U}}{\partial t} + \nabla \cdot \left( \vec{U} \otimes \vec{U}\right) - \nabla \cdot \nu \left( \nabla \vec{U} + (\nabla \vec{U})^T\right) = - \frac{1}{\rho} \nabla p,
 \]

### 2021-07-07 2021/Entropy/entropy.tex
some corrections if algorithm description

@@
  Step 1. Перенумеровать нижним индексом (начиная с 0) точки $x^i,\:0\leq i\leq k$, проведенных испытаний в порядке возрастания координаты, т.е.
-\[
+\begin{equation}\label{xt}
 0=x_0<x_1<\ldots <x_{k}=1.
-\] 
+\end{equation} 
 Сопоставить точкам $x_i, \; 0\leq i\leq k$, вычисленные в них значения целевой функции $z_i=\varphi(x_i), \; 0\leq i\leq k$.
@@
 Выделение областей притяжения локальных минимумов с использованием модели (\ref{dtree}) можно организовать следующим образом.
-%Здесь - описание алгоритма выделения локальных минимумов, с иллюстрацией на функции Шекеля.
 
-Пусть $x^{k+1}$ -- текущая точка испытания. Для данной точки ищется номер $j$, такой что $x^{k+1} \in D_j$ 
+Пусть $x^{k+1}$ -- точка текущего испытания. Для данной точки ищется номер $j$, такой что $x^{k+1} \in D_j$. 
 Далее сравниваются значения $c_j$, соответствующие соседним подобластям. Если выполняется одно из условий
-\begin{gather}\label{LocalPointC}
-c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3} \leq c_{j+4}, \; j=1, \\ \nonumber
-c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3}, \; j=2,  \\ \nonumber
-c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2}, \; 3 \leq j \leq J-2;\\
-c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1}, \; j=J-1, \\ \nonumber
-c_{j-4} \geq c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j}, \; j=J, \nonumber
-\end{gather}
-то область $D_j$ считается областью притяжения локального минимума. В ней можно запускать локальный поиск.
+\begin{gather*}\label{LocalPointC}
+c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3} \leq c_{j+4}, \; j=1, \\ 
+c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3}, \; j=2,  \\ 
+c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2}, \; 3 \leq j \leq J-2; \\ 
+c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1}, \; j=J-1, \\ 
+c_{j-4} \geq c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j}, \; j=J, 
+\end{gather*}
+то область $D_j$ считается областью притяжения локального минимума. В ней можно запускать локальный поиск, а впоследствии -- исключить данную подобласть из глобального поиска.
+В качестве локального алгоритма может использоваться любой из локальных методов нулевого порядка, например,  golden section search or parabolic interpolation \cite{Press}.
 
-%\begin{equation}\label{LocalPointC}
-%c_{j-t_1} > ... > c_{j-1} > c_{j} < c_{j+1} < c_{j+t_2}
-%\end{equation}
-%\begin{equation}\label{LocalPointT}
- %t_1 + t_2 + 1  = T,
-%\end{equation}
-%где $T$ -- параметр метода, экспериментально было выбрано значение 5.
-%Тогда область $D_j$ является областью притяжения локального минимума.
+Для того, чтобы модифицировать алгоритм глобального поиска из subsection \ref{CoreGSA} с целью исключения областей притяжения локальных минимумов, сопоставим каждой точке испытания $x^i$, получаемой при работе алгоритма, дополнительный признак $q^i \in \{0,1,2\}$, который будет характеризовать свойства данной точки.
+Значение $q^i=0$ присваивается по умолчанию и означает что точка $x^i$ получена в результате работы правила (\ref{xk1}) алгоритма глобального поиска.
+Значение $q^i=1$ присваивается точкам, полученным в результате работы локального метода, при этом значение $q^i=2$ соответствует точке локального минимума, найденной в результате работы локального поиска.
 
-Для того, чтобы использовать выделение локальных минимумов в алгоритме глобального поиска, описанном в subsection \ref{CoreGSA}, всем точкам получаемым в ходе работы алгоритма присваивается метка q, со значениями global - по умолчанию, означает что точка получена в результате описанного алгоритма глобального поиска, local - точка получена в результате работы локального метода, localMin - точка является найденным локальным минимумом.
-Индексы метки соответствуют индексам координаты $x$, т.е.
-у точки  $x^{k+1}$ метка $q^{k+1}$ и так далее.
+Напомним, что верхний индекс соответствует номеру итерации, на которой было проведено испытание в данной точке, а нижний индекс -- номеру точки в ряду (\ref{xt}).
 
-Также алгоритм необходимо дополнить следующими шагами.
+Опишем теперь модифицированный алгоритм глобального поиска (GSA), использующий деревья решений для выделения и исключения областей притяжения локальных минимумов; будем в дальнейшем ссылаться на данный алгоритм как на GSA-DT.
 
-Step 6.
-Если хотя бы одна из точек $x_{t-1}$ или $x_t$ имеет метку local или localMin, то точка $x^{k+1}$ исключается из поисковой информации и переходим к проверке критерия остановки.
-Иначе переходим на Step 7.
+Шаги 1--5 алгоритма GSA-DT полностью повторяют шаги 1--5 .
+
+Step 6. Если $q_{t-1} \in \{1,2\}$ или $q_t \in \{1,2\}$, то перейти к проверке критерия остановки. Иначе перейти на Step 7.
 
@@
 Step 8.
-Для точки $x^{k+1}$ текущего испытания найти $j$ такой, что $x^{k+1} \in D_j$ и проверить выполнимость условия (\ref{LocalPointC}).
-Если условие (\ref{LocalPointC}) выполняется, запустить локальный поиск из точки $x^{k+1}$.
-Все $N_{local}$ точек полученных локальным методом сохраняются и получают номера $x^{k+2}, ..., x^{k+N_{local}+1}$  и используются на последующих итерациях алгоритма.
-Также эти точки получают метку local, $q^{i}=local, i = k+2, ...
-, k+N_{local}+1 $.
-Точка соответствующая найденному локальному минимуму получает метку localMin.
+Для точки $x^{k+1}$ текущего испытания найти номер $j$ такой, что $x^{k+1} \in D_j$ и проверить выполнимость условия (\ref{LocalPointC}).
+Если условие (\ref{LocalPointC}) выполняется, запустить в области $D_j$ локальный поиск из точки $x^{k+1}$.
+Результаты всех испытаний, выполненных в процессе работы локального поиска в точках $x^{k+2}, ...,x^{k+k_{local}}$, сохраняются в информационной базе алгоритма и используются на последующих итерациях.
+Все эти точки получают признак $q^i=1$, $i = k+2, ... , k+k_{local}$. Точке, соответствующей найденному локальному минимуму, присваивается признак равный 2.
 
-В качестве локального алгоритма на шаге 7 может использоваться любой из локальных методов нулевого порядка, например,  golden section search or parabolic interpolation \cite{Press}.
+Критерий остановки модифицированного алгоритма будет выглядеть следующим образом. 
+Алгоритм прекращает свою работу при выполнении одного из условий:
 
-Критерий остановки будет выглядеть следующим образом:
- Алгоритм прекращает свою работу при выполнении одного из условий
+\begin{itemize}
+	\item $|x_{t} - x_{t-1}|<\epsilon$, где $t$ из (\ref{MaxR}), а $\epsilon>0$ есть заданная точность;
+	\item $|x^{k+1} - x_{t-1}|<\epsilon$ и $q_{t-1} = 2$,
+	\item $|x^{k+1}-x_{t}|<\epsilon$ и $x_{t} = 2$.
+\end{itemize}
 
-1. $\Delta_t<\epsilon$; здесь $t$ из (\ref{MaxR}), а $\epsilon>0$ есть заданная точность.
-
-2. $|x^{k+1}_i-x_{t-1}|<\epsilon$ и $x_{t-1}$ имеет метку localMin.
-
-3. $|x^{k+1}_i-x_{t}|<\epsilon$ и $x_{t}$ имеет метку localMin.
-
-%Нужно еще описать исключение интервалов из глоабльно поиска, и также - новый критерий остановки.
+Проиллюстрируем работу алгоритма при минимизации двух конкретных многоэкстремальных функций, которые задаются в соответствии с формулами (\ref{shekel}) и (\ref{hill}). 
+%далее нужны иллюстрации
 
@@
 
-В работе \cite{Gergel2019} описан GCGen-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: размерностью задачи, точкой глобального минимума, значением функции в ней и т.п.
-Испытания будем проводить на функциях Шекеля и Хилла-Гибсона.
+Традиционный подход к оценке эффективности методов глобальной оптимизации основан на численном решении этими методами серии задач из некоторого класса. 
+При этом подразумевается, что очередная решаемая задача генерируется посредством некоторого алгоритма.
+Типичными примерами таких классов тестовых функций являются функции  Шекеля и Хилла. 
+Первый из них (обозначит его $F_{SH}$) основан на формуле 
+\begin{equation}\label{shekel}
+  \varphi(x)=-\sum_{j=1}^{10}\frac{1}{(K_j(x-A_j)^2+C_j)},\;  x\in[0,10],
+\end{equation}
+где параметры $1\le K_j\le 3,\: 0 < A_j,\: C_j < 10, \;$ являются независимыми случайными величинами, равномерно распределенными в указанных интервалах.
+Следующий генератор (обозначим его $F_{HL}$) определяется выражением
+\begin{equation}\label{hill}
+  \varphi(x)=\sum_{j=1}^{14}(A_j\sin(2j\pi x) + B_j\cos(2j\pi x)),\: x\in[0,1]
+\end{equation}
+где значения параметров  $A_j,\: B_j,\: 1 \le j \le 14$, независимо и равномерно распределены в интервале $[-1 ,1 ]$.
+
+Проведем сравнение базового global search algorithm (GSA) и его модификации, использующей деревья решений (GSA-DT), c известным алгоритмом глобальной оптимизации DIRECT \cite{Jones2009}.
+Сравнение методов будем проводить при решении 1000 задач из классов $F_{SH}$ и $F_{HL}$.
+
+Задачу будем считать решенной корректно в случае, если после остановки метода по точности (т.е.
+когда длина текущего поискового интервала станет меньше $\epsilon$) текущая оценка оптимума $x_k^*$ будет лежать в $\epsilon$-окрестности известного решения задачи $x^*$, т.е.
+будет выполняться условие $|x^*-x_k^*| \leq \epsilon$.
+
+В таблицах \ref{table:average_Shekel} и \ref{table:average_Hill} приведено число поисковых испытаний, которые в среднем требовалось для минимизации функций Шекеля и Хилла с разной точностью поиска.
+При этом в скобках в таблице указано число нерешенных задач.
+
+Результаты экспериментов показывают, что при грубой точности решения задачи все методы показывают схожие результаты по числу trials, тогда как при высокой точности решения  алгоритму GSA-DT требуется в 2 раза меньше trials, чем его прототипу.
+Одновременно GSA-DT превосходит метод DIRECT как по среднему числу search trials, так и по числу корректно решенных задач.
 
 \begin{specialtable}[H] 
-	\caption{Average number of iterations, Shekel problem}\label{table:average_Shekel}
+	\caption{Среднее число испытаний при минимизации Shekel test functions}\label{table:average_Shekel}
 	\center
@@
 \midrule													
-DiRect                                              & 64,22                                   &                                         &                                        &                                        &                                       \\
-Базовый АГП                                         & 106                                     & 64                                      & 53                                     & 37                                     & 31                                    \\
-АГП с деревом решений & 49                                      & 45                                      & 43                                     & 38                                     & 35                                   \\
+DIRECT         & 64   &     &     &     &      \\
+GSA            & 106  & 64  & 53  & 37  & 31   \\
+GSA-DT         & 49   & 45  & 43  & 38  & 35   \\
 
@@
 \begin{specialtable}[H] 
-	\caption{Average number of iterations Hill problem}\label{table:average_Hill}
+	\caption{Среднее число испытаний при минимизации Hill test functions}\label{table:average_Hill}
 	\center
@@
 \midrule					  
-DiRect                & 65,74 (91)                              &                                         &                                        &                                        &                                       \\
-Базовый АГП           & 130                                     & 87                                      & 75                                     & 52                                     & 43                                    \\
-АГП с деревом решений & 64                                      & 60                                      & 59                                     & 55                                     & 50                                    \\
+DIRECT                & 65,74 (91)      &                 &                &                &               \\
+GSA                   & 130             & 87              & 75             & 52             & 43            \\
+GSA-DT                & 64              & 60              & 59             & 55             & 50            \\
 \bottomrule
@@
 
-В таблицах \ref{table:average_Shekel} и \ref{table:average_Hill} приведено среднее число итераций полученное при решении серии из 100 задач Шекеля и Хилла, для алгоритмов Direct, АГП и АГП с использованием локального метода запущенного по результатам работы дерева решений.
-Здесь и далее в скобках указано сколько задач решилось их 100.
-
-In the paper \cite{Gaviano2003} a GKLS generator is described that allows one to generate multiextremal optimization problems with previously known properties: the number of local minima, the size of their regions of attraction, the global minimum point, the value of the function in it, etc.
+Следующую серию экспериментов проведем при решении многомерных задач. 
+Известным генератором тестовых задач многоэкстремальной оптимизации является генератор GLKS \cite{Gaviano2003}.
+С его помощью можно порождать задачи оптимизации с заданными свойствами: числом локальных экстремумов, их областями притяжения, точкой глобального минимума и значением целевой функции в ней, и т.д.
 

### 2021-07-09 2021/KMC/abstract.tex
Статья на конференцию математичесих центров, по решению химической задачи

\documentclass[12pt]{article}

\usepackage[cp1251]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{epsfig,graphicx}

\usepackage{graphicx}
\usepackage{marvosym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cite}

\usepackage{url}
\usepackage{hyperref}
\def\UrlFont{\rmfamily}

\def\orcidID#1{\unskip$^{[#1]}$}
\def\letter{$^{\textrm{(\Letter)}}$}

\oddsidemargin 0cm
\topmargin 0cm
\textwidth=16cm
\textheight=23cm
\thispagestyle{empty}

\begin{document}
\renewcommand{\refname}{}%<----------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large\bf
Parallel Global Search Algorithm for Optimization of the Kinetic Parameters of Chemical Reactions\bigskip\\} %<== put the title here
Irek Gubaydullin$^{1,2}$,
Leniza Enikeeva$^{2,3}$,
Konstantin Barkalov$^4$,
Ilya Lebedev$^4$\medskip\\       %<== put the authors here
{\em $^1$Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia, $^2$Ufa State Petroleum Technological University, Ufa, Russia, $^3$Novosibirsk State University, Novosibirsk, Russia, $^4$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\}             %<== put affiliation(s) here
\bigskip
\end{center}

% <==Put your text here.

Одним из направлений лаборатории математической химии Института нефтехимии и катализа УФИЦ РАН является исследование химических реакций методами математического моделирования. Кинетическая модель является основой математического моделирования химических реакций и исходным уровнем при математическом моделировании химических реакторов. С математической точки зрения обратная задача химической кинетики представляет собой задачу оптимизации, поэтому актуальной является задача выбора метода оптимизации.  В данной работе рассматривается решение этой задачи параллельным алгоритмом глобальной оптимизации. Теоретическим базисом данного алгоритма является информационно-статистический подход, детально изложенный в \cite{Strongin2000}. Разработанный алгоритм основан на редукции исходной многомерной задачи к эквивалентной ей одномерной задаче с последующим ее решением эффективными методами оптимизации функций одной переменной. Вопросы распараллеливания алгоритма на разных архитектурах рассмотрены в \cite{Barkalov2016}.

Исследуемый процесс представляет собой преобразование пропана в газ, богатый метаном, на никелевом катализаторе, который является промышленно важным химическим процессом \cite{Uskov2019, Uskov2020}. Экспериментальные данные по предварительному риформингу пропана были получены из работы \ cite {Uskov2020}. Схема реакции состоит из двух реакций: паровая конверсияпропана и метанирование CO$_2$ \cite{Uskov2017}
\begin{gather}
C_3H_8 + 6H_2O \rightarrow 10H_2 + 3CO_2 \label{reac_ref}\\
CO_2 + 4H_2 \rightleftharpoons CH_4 + 2H_2O \label{reac_met},
\end{gather}

Математическая модель задач химической кинетики представляет собой систему дифференциальных уравнений, которая описывает изменения концентраций веществ во времени в соответствии со скоростью стадий реакции. Решение системы обыкновенных дифференциальных уравнений представляет собой вектор рассчитанных концентраций компонентов реакции, которые затем сравниваются с экспериментальными данными.
Математическая задача - минимизировать функционал отклонения расчетных значений от экспериментальных. Целевая функция определяется как сумма абсолютных отклонений расчетных и экспериментальных концентраций:

\begin{equation}\label{func}
F = \sum\limits_{i=1}^M \sum\limits_{j=1}^N \left| x_{ij}^{calc} - x_{ij}^{exp} \right| \rightarrow \min,
\end{equation}

где $ x_ {ij} ^ {calc} $ и $ x_ {ij} ^ {exp} $ - расчетные и экспериментальные значения концентраций компонентов; $ M $ - количество точек измерения; $ N $ - количество веществ, участвующих в реакции.

Таким образом, в результате решения поставленной задачи химической кинетики параллельным методом глобального поиска были рассчитаны кинетические параметры процесса предриформинга пропана на Ni катализаторе.

Для оценки эффективности распараллеливания алгоритм глобального поиска запускался в последовательном и паралелльном режимах, максимальное число задействованных вычислительных узлов составляло 20, соответственно, максимальное число задействованных процессоров (P) было 40. В таблице \ref{table:Acceleration} приведено ускорение полученное при решение задачи.

\begin{table}[]
	\caption{Ускорение на P процессах }\label{table:Acceleration}
	\center
\begin{tabular}{ccc}

         \textbf{P=10} & \textbf{P=20} & \textbf{P=40}  \\

          5,74  &  8,58  &  15,63  \\

\end{tabular}
\end{table}

Вычислительные эксперименты проводились на кластере «Лобачевский», Узел кластера состоит из 2-х Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM.

\begin{thebibliography}{10}

\bibitem{Strongin2000} Strongin R.G., Sergeyev Y.D. Global optimization with non-convex constraints. Sequential and parallel algorithms, Kluwer Academic Publishers, Dordrecht, {2000}

\bibitem{Barkalov2016} Barkalov, K. and Lebedev, I. Solving multidimensional global optimization problems using graphics accelerators. 
Communications in Computer and Information Science, volume 687, p. 224-235. 2016

\bibitem{Uskov2019} Uskov, S.I. and Potemkin, D.I. and Kamboj, N. and Snytnikov, P.V. and  Pakharukova,V.P. and Enikeeva, L.V. and Gubaydullin, I.M. and Hussainova, I. Fibrous alumina-based Ni-MOx (M = Mg, Cr, Ce) catalysts for propane pre-reforming. Materials Letters, 	volume = 257, 	p. 126741, 	{2019}

\bibitem{Uskov2020} Uskov, S.I. and Potemkin, D.I. and Enikeeva, L.V. and Snytnikov, P.V. and Gubaydullin, I.M. and Sobyanin, V.A. Propane Pre-Reforming into Methane-Rich Gas over Ni Catalyst: Experiment and Kinetics Elucidation via Genetic Algorithm. Energies, volume = 13, {2020}

\bibitem{Uskov2017} Uskov, S.I. and  Enikeeva, L.V. and Potemkin, D. I. and Belyaev, V. D. and Snytnikov, P.V. and Gubaidullin, I.M. and  Kirillov, V. A. and  Sobyanin, V. A. Kinetics of low-temperature steam reforming of propane in a methane excess on a Ni-based catalyst.Catalysis in Industry. volume = 9, p. 104 -- 109, {2017}

\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2021-07-09 2021/KMC/PGSA.tex
Добавлена вторая версия

--- 2021-07-09 2021/KMC/abstract.tex
@@
-\documentclass[12pt]{article}
-
-\usepackage[cp1251]{inputenc}
-\usepackage[english,russian]{babel}
-\usepackage{epsfig,graphicx}
+\documentclass[11pt, oneside, a4paper]{article}
+\usepackage[utf8]{inputenc}
+%\usepackage[cp1251]{inputenc} % кодировка
+\usepackage[english, russian]{babel} % Русские и английские переносы
+\usepackage{graphicx}          % для включения графических изображений
+\usepackage{cite}              % для корректного оформления литературы
+\usepackage{pavt-ru}
 
@@
 \usepackage{amssymb}
-\usepackage{cite}
 
@@
 
-\oddsidemargin 0cm
-\topmargin 0cm
-\textwidth=16cm
-\textheight=23cm
-\thispagestyle{empty}
+\begin{document}
 
-\begin{document}
-\renewcommand{\refname}{}%<----------
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \title - название статьи
+% \authors - список авторов
 
-\begin{center}
-{\large\bf
-Parallel Global Search Algorithm for Optimization of the Kinetic Parameters of Chemical Reactions\bigskip\\} %<== put the title here
-Irek Gubaydullin$^{1,2}$,
-Leniza Enikeeva$^{2,3}$,
-Konstantin Barkalov$^4$,
-Ilya Lebedev$^4$\medskip\\       %<== put the authors here
-{\em $^1$Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia, $^2$Ufa State Petroleum Technological University, Ufa, Russia, $^3$Novosibirsk State University, Novosibirsk, Russia, $^4$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\}             %<== put affiliation(s) here
-\bigskip
-\end{center}
+\title{Parallel Global Search Algorithm }
 
-% <==Put your text here.
+\authors{К.А.~Баркалов, И.Г.~Лебедев}
+\organizations{Нижегородский государственный университет им. Н.И. Лобачевского}
 
-Одним из направлений лаборатории математической химии Института нефтехимии и катализа УФИЦ РАН является исследование химических реакций методами математического моделирования.
-Кинетическая модель является основой математического моделирования химических реакций и исходным уровнем при математическом моделировании химических реакторов.
+Кинетическая модель является одной из основ математического моделирования химических реакций и исходным уровнем при математическом моделировании химических реакторов.
 С математической точки зрения обратная задача химической кинетики представляет собой задачу оптимизации, поэтому актуальной является задача выбора метода оптимизации.
 В данной работе рассматривается решение этой задачи параллельным алгоритмом глобальной оптимизации.
-Теоретическим базисом данного алгоритма является информационно-статистический подход, детально изложенный в \cite{Strongin2000}.
 Разработанный алгоритм основан на редукции исходной многомерной задачи к эквивалентной ей одномерной задаче с последующим ее решением эффективными методами оптимизации функций одной переменной.
-Вопросы распараллеливания алгоритма на разных архитектурах рассмотрены в \cite{Barkalov2016}.
+Вопросы распараллеливания алгоритма на разных архитектурах рассмотрены в \cite{Strongin13, Barkalov2016}.
 
-Исследуемый процесс представляет собой преобразование пропана в газ, богатый метаном, на никелевом катализаторе, который является промышленно важным химическим процессом \cite{Uskov2019, Uskov2020}.
-Экспериментальные данные по предварительному риформингу пропана были получены из работы \ cite {Uskov2020}.
-Схема реакции состоит из двух реакций: паровая конверсияпропана и метанирование CO$_2$ \cite{Uskov2017}
-\begin{gather}
-C_3H_8 + 6H_2O \rightarrow 10H_2 + 3CO_2 \label{reac_ref}\\
-CO_2 + 4H_2 \rightleftharpoons CH_4 + 2H_2O \label{reac_met},
-\end{gather}
+Исследуемый процесс представляет собой преобразование пропана в газ, богатый метаном, на никелевом катализаторе, который является промышленно важным химическим процессом.
+Экспериментальные данные по предварительному риформингу пропана были получены в работе \cite {Uskov2020}.
+Схема реакции состоит из двух реакций: паровая конверсияпропана и метанирование CO$_2$
 
 Математическая модель задач химической кинетики представляет собой систему дифференциальных уравнений, которая описывает изменения концентраций веществ во времени в соответствии со скоростью стадий реакции.
-Решение системы обыкновенных дифференциальных уравнений представляет собой вектор рассчитанных концентраций компонентов реакции, которые затем сравниваются с экспериментальными данными.
-Математическая задача - минимизировать функционал отклонения расчетных значений от экспериментальных.
-Целевая функция определяется как сумма абсолютных отклонений расчетных и экспериментальных концентраций:
+Согласно этой моделе можно сформировать оптимизационную задачу, в которой целевая функция определяется как сумма абсолютных отклонений расчетных и экспериментальных концентраций:
 
@@
 
-Таким образом, в результате решения поставленной задачи химической кинетики параллельным методом глобального поиска были рассчитаны кинетические параметры процесса предриформинга пропана на Ni катализаторе.
+В результате решения поставленной задачи химической кинетики параллельным методом глобального поиска были рассчитаны кинетические параметры процесса предриформинга пропана на Ni катализаторе.
 
-Для оценки эффективности распараллеливания алгоритм глобального поиска запускался в последовательном и паралелльном режимах, максимальное число задействованных вычислительных узлов составляло 20, соответственно, максимальное число задействованных процессоров (P) было 40.
-В таблице \ref{table:Acceleration} приведено ускорение полученное при решение задачи.
-
-\begin{table}[]
-	\caption{Ускорение на P процессах }\label{table:Acceleration}
-	\center
-\begin{tabular}{ccc}
-
-         \textbf{P=10} & \textbf{P=20} & \textbf{P=40}  \\
-
-          5,74  &  8,58  &  15,63  \\
-
-\end{tabular}
-\end{table}
+Алгоритм глобального поиска запускался в последовательном и паралелльном режимах, максимальное число задействованных процессоров (P) состовляло 40, и ускорение составило соответственно 15,63.
 
@@
 
-\begin{thebibliography}{10}
+\begin{biblio}
 
-\bibitem{Strongin2000} Strongin R.G., Sergeyev Y.D. Global optimization with non-convex constraints. Sequential and parallel algorithms, Kluwer Academic Publishers, Dordrecht, {2000}
+\bibitem{Strongin13}
+Стронгин~Р.Г., Гергель~В.П., Гришагин~В.A., Баркалов~К.А. Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013. 280~с.
 
@@
 Communications in Computer and Information Science, volume 687, p. 224-235. 2016
-
-\bibitem{Uskov2019} Uskov, S.I.
-and Potemkin, D.I.
-and Kamboj, N.
-and Snytnikov, P.V.
-and  Pakharukova,V.P.
-and Enikeeva, L.V.
-and Gubaydullin, I.M.
-and Hussainova, I.
-Fibrous alumina-based Ni-MOx (M = Mg, Cr, Ce) catalysts for propane pre-reforming.
-Materials Letters, 	volume = 257, 	p.
-126741, 	{2019}
 
@@
 Propane Pre-Reforming into Methane-Rich Gas over Ni Catalyst: Experiment and Kinetics Elucidation via Genetic Algorithm.
-Energies, volume = 13, {2020}
+Energies, volume  13, {2020}
 
-\bibitem{Uskov2017} Uskov, S.I.
-and  Enikeeva, L.V.
-and Potemkin, D.
-I.
-and Belyaev, V.
-D.
-and Snytnikov, P.V.
-and Gubaidullin, I.M.
-and  Kirillov, V.
-A.
-and  Sobyanin, V.
-A.
-Kinetics of low-temperature steam reforming of propane in a methane excess on a Ni-based catalyst.Catalysis in Industry.
-volume = 9, p.
-104 -- 109, {2017}
-
-\end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\end{biblio}
 \end{document}

### 2021-07-09 2021/KMC/PGSA.tex
Обновленный вариант текста

@@
 
-\title{Parallel Global Search Algorithm }
+\title{Идентификация кинетических параметров химической реакции с помощью параллельного алгоритма глобального поиска }
 
@@
 
-Исследуемый процесс представляет собой преобразование пропана в газ, богатый метаном, на никелевом катализаторе, который является промышленно важным химическим процессом.
-Экспериментальные данные по предварительному риформингу пропана были получены в работе \cite {Uskov2020}.
-Схема реакции состоит из двух реакций: паровая конверсияпропана и метанирование CO$_2$
-
+В данной работе исследуется одна из химических реакций, экспериментальные данные и ее описание рассмотрены в работе \cite {Uskov2020}.
 Математическая модель задач химической кинетики представляет собой систему дифференциальных уравнений, которая описывает изменения концентраций веществ во времени в соответствии со скоростью стадий реакции.
@@
 
-\begin{equation}\label{func}
+\begin{displaymath}\label{func}
 F = \sum\limits_{i=1}^M \sum\limits_{j=1}^N \left| x_{ij}^{calc} - x_{ij}^{exp} \right| \rightarrow \min,
-\end{equation}
-
-где $ x_ {ij} ^ {calc} $ и $ x_ {ij} ^ {exp} $ - расчетные и экспериментальные значения концентраций компонентов; $ M $ - количество точек измерения; $ N $ - количество веществ, участвующих в реакции.
+\end{displaymath}
+где $ x_ {ij} ^ {calc} $ и $ x_ {ij} ^ {exp} $ -- расчетные и экспериментальные значения концентраций компонентов; $ M $ -- количество точек измерения; $ N $ -- количество веществ, участвующих в реакции.
 
@@
 
-Алгоритм глобального поиска запускался в последовательном и паралелльном режимах, максимальное число задействованных процессоров (P) состовляло 40, и ускорение составило соответственно 15,63.
-
-Вычислительные эксперименты проводились на кластере «Лобачевский», Узел кластера состоит из 2-х Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM.
+Вычислительные эксперименты проводились на кластере «Лобачевский», Узел кластера состоит из 2-х Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM.
+Алгоритм глобального поиска запускался в последовательном и паралелльном режимах, максимальное число задействованных процессоров состовляло 40, и ускорение составило соответственно 15.63.
 

### 2021-07-09 2021/KMC/PGSA.tex
Поправлен текст

@@
 
-Кинетическая модель является одной из основ математического моделирования химических реакций и исходным уровнем при математическом моделировании химических реакторов.
-С математической точки зрения обратная задача химической кинетики представляет собой задачу оптимизации, поэтому актуальной является задача выбора метода оптимизации.
+Основным способом математического моделирования химических реакций является построение кинетической модели.
+Также эта модель необходима при моделировании химических реакторов.
+При разработке кинетической модели многостадийной химической реакции решается обратная задача химической кинетики, которая представляет собой задачу оптимизации.
 В данной работе рассматривается решение этой задачи параллельным алгоритмом глобальной оптимизации.

### 2021-07-09 2021/KMC/PGSA.tex
udpate first paragraph and references

@@
 Основным способом математического моделирования химических реакций является построение кинетической модели.
-Также эта модель необходима при моделировании химических реакторов.
-При разработке кинетической модели многостадийной химической реакции решается обратная задача химической кинетики, которая представляет собой задачу оптимизации.
+При разработке кинетической модели многостадийной химической реакции решается обратная задача химической кинетики, которая представляет собой задачу глобальной оптимизации с целевой функцией вида ``черный ящик''.
 В данной работе рассматривается решение этой задачи параллельным алгоритмом глобальной оптимизации.
@@
 
-В данной работе исследуется одна из химических реакций, экспериментальные данные и ее описание рассмотрены в работе \cite {Uskov2020}.
+Исследуется одна из химических реакций, экспериментальные данные и ее описание рассмотрены в работе \cite {Uskov2020}.
 Математическая модель задач химической кинетики представляет собой систему дифференциальных уравнений, которая описывает изменения концентраций веществ во времени в соответствии со скоростью стадий реакции.
@@
 
-\bibitem{Barkalov2016} Barkalov, K. and Lebedev, I. Solving multidimensional global optimization problems using graphics accelerators. 
-Communications in Computer and Information Science, volume 687, p. 224-235. 2016
+\bibitem{Barkalov2016} Barkalov K., Lebedev I.
+Solving multidimensional global optimization problems using graphics accelerators // Communications in Computer and Information Science.
+2016.
+vol.
+687, pp.
+224-235.
 
-\bibitem{Uskov2020} Uskov, S.I.
-and Potemkin, D.I.
-and Enikeeva, L.V.
-and Snytnikov, P.V.
-and Gubaydullin, I.M.
-and Sobyanin, V.A.
-Propane Pre-Reforming into Methane-Rich Gas over Ni Catalyst: Experiment and Kinetics Elucidation via Genetic Algorithm.
-Energies, volume  13, {2020}
+\bibitem{Uskov2020} Uskov S.I., Potemkin D.I., Enikeeva L.V., Snytnikov P.V., Gubaydullin I.M., Sobyanin V.A.
+Propane pre-reforming into methane-rich gas over Ni catalyst: experiment and kinetics elucidation via genetic algorithm.
+Energies.
+2020.
+vol.
+13, art.no 3393.
 

### 2021-07-10 2021/Entropy/entropy.tex
Небольшая правка и добавлена информация по direct
virst version of figures
Данные по запуску GKLS 2d на direct

@@
 	\item $|x^{k+1} - x_{t-1}|<\epsilon$ и $q_{t-1} = 2$,
-	\item $|x^{k+1}-x_{t}|<\epsilon$ и $x_{t} = 2$.
+	\item $|x^{k+1}-x_{t}|<\epsilon$ и $q_{t} = 2$.
 \end{itemize}
@@
 %далее нужны иллюстрации
+\begin{figure}[H]
+\label{fig1}
+\includegraphics[width=1.0\linewidth]{hill.png}
+\caption{This is a figure. A caption on a single line should be centered.}
+\end{figure}
+
+\begin{figure}[H]
+\label{fig2}
+\includegraphics[width=1.0\linewidth]{shekel.png}
+\caption{This is a figure. A caption on a single line should be centered.}
+\end{figure}
 
@@
 \midrule													
-DIRECT         & 64   &     &     &     &      \\
+DIRECT         & 64(99)   &  -   &  34(94)   &  -   &  20(83)    \\
 GSA            & 106  & 64  & 53  & 37  & 31   \\
@@
 \midrule					  
-DIRECT                & 65,74 (91)      &                 &                &                &               \\
+DIRECT                & 66(88) &  -  &  36(69)  & -  &   20(49)  \\
 GSA                   & 130             & 87              & 75             & 52             & 43            \\
@@
 					  \midrule
-DiRect                &                                        & 80,04 (41)                            &                                        &                                       &                &                                       \\
+DiRect                &   39(0)  & 29,92
+ (6)                            &                                        &                                       &                &                                       \\
 Базовый АГП           & 1820                                   & 841                                   & 53251,32                               & 8905,03                               & 665013,15      & 135972,78 (99)                        \\

### 2021-07-10 2021/Entropy/entropy.tex
Дробная часть

@@
 					  \midrule
-DiRect                &   39(0)  & 29,92
+DiRect                &   39,84(0)  & 29,92
  (6)                            &                                        &                                       &                &                                       \\

### 2021-07-27 2021/Entropy/entropy.tex
Update 1d illustrations

@@
 Step 7.
-Построить дерево решений на основе результатов выполненных испытаний, и получить соответствующую кусочно линейную аппроксимацию, которая каждой подобласти $D_1, D_2, ..., D_J$ сопоставляет значение $с_1, с_2, ..., с_J$.
+Построить дерево решений на основе результатов выполненных испытаний, и получить соответствующую кусочно линейную аппроксимацию, которая каждой подобласти $D_1, D_2, ..., D_J$ сопоставляет значение $c_1, c_2, ..., c_J$.
 
@@
 
-Проиллюстрируем работу алгоритма при минимизации двух конкретных многоэкстремальных функций, которые задаются в соответствии с формулами (\ref{shekel}) и (\ref{hill}). 
+Рисунки \ref{fig1} и figure \ref{fig2} иллюстрируют работу алгоритма при минимизации конкретной многоэкстремальной функции, которая задаются в соответствии с формулой (\ref{hill}).
+При запуске алгоритма использовался параметр $r=2.2$ из (\ref{R}) и значение $\epsilon = 10^{-3}$ в критерии остановки метода.
+
+На рисунке \ref{fig1} изображен график objective function и точки 71 поискового испытания, потребовавшихся GSA для решения задачи с указанной точностью.
+Наглядно видна концентрация точек в окрестностях локальных минимумов задачи.
+Рисунок \ref{fig2} соответствуют работе алгоритма GSA-DT.
+Кроме графика целевой функици на нем изображена кусочно-постоянная аппроксимация вида (\ref{dtree}), построенная на финальной стадии поиска.
+Черные точки на графике соответствуют фазе глобального поиска, зеленые точки соответствуют работе локального метода.
+Всего методу GSA-DT потребовалось 49 испытаний для решения задачи; накопление точек испытаний в окрестностях локальных минимумов отсутствует.
+
 %далее нужны иллюстрации
@@
 \label{fig1}
-\includegraphics[width=1.0\linewidth]{hill.png}
-\caption{This is a figure. A caption on a single line should be centered.}
+\includegraphics[width=1.0\linewidth]{HillAGP90.png}
+\caption{Минимизация функции вида (\ref{hill}) с помощью GSA.}
 \end{figure}
@@
 \label{fig2}
-\includegraphics[width=1.0\linewidth]{shekel.png}
-\caption{This is a figure. A caption on a single line should be centered.}
+\includegraphics[width=1.0\linewidth]{HillTree90.png}
+\caption{Минимизация функции вида (\ref{hill}) с помощью GSA-DT.}
 \end{figure}
@@
 Одновременно GSA-DT превосходит метод DIRECT как по среднему числу search trials, так и по числу корректно решенных задач.
+В частности, если использовать точность $\epsilon = 10^{-2}$, то метод DIRECT останавливается слишком рано, и не находит глобального решения во многих задачах.
 
@@
 	\center
-\begin{tabular}{cccccc}
+\begin{tabular}{cccc}
 \toprule
-        & \textbf{0,0001} & \textbf{0,0005} & \textbf{0,001} & \textbf{0,005} & \textbf{0,01} \\
+        & \textbf{$\epsilon = 10^{-4}$} & \textbf{$\epsilon = 10^{-3}$} & \textbf{$\epsilon = 10^{-2}$} \\
 \midrule													
-DIRECT         & 64(99)   &  -   &  34(94)   &  -   &  20(83)    \\
-GSA            & 106  & 64  & 53  & 37  & 31   \\
-GSA-DT         & 49   & 45  & 43  & 38  & 35   \\
+DIRECT         & 64(1) &  34(6)   & 20(17)    \\
+GSA            & 106  & 53  &  31   \\ 
+GSA-DT         & 49   & 43  &  35   \\
 
@@
 	\center
-\begin{tabular}{cccccc}
+\begin{tabular}{cccc}
 \toprule
-                      & \textbf{0,0001} & \textbf{0,0005} & \textbf{0,001} & \textbf{0,005} & \textbf{0,01} \\
+        & \textbf{$\epsilon = 10^{-4}$} & \textbf{$\epsilon = 10^{-3}$} & \textbf{$\epsilon = 10^{-2}$} \\
 \midrule					  
-DIRECT                & 66(88) &  -  &  36(69)  & -  &   20(49)  \\
-GSA                   & 130             & 87              & 75             & 52             & 43            \\
-GSA-DT                & 64              & 60              & 59             & 55             & 50            \\
+DIRECT                & 66(12) & 36(31)  & 20(51)  \\
+GSA                   & 130    & 75      & 43      \\
+GSA-DT                & 64     & 59      & 50      \\
 \bottomrule
@@
 В таблицах \ref{table:average_GKLS} приведено среднее число итераций полученное на  задачах GKLS, для алгоритмов Direct, АГП, АГП с адаптивной схемой и АГП с адаптивной схемой с использованием локального метода запущенного по результатам работы дерева решений.
-
-\section{Conclusions}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-
-Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
-Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
-Please disclose at the submission stage any restrictions on the availability of materials or information.
-New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
-
-Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers.
-If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
-They must be provided prior to publication.
-
-Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
-\begin{quote}
-This is an example of a quote.
-\end{quote}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-
-This section may be divided by subheadings.
-It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
-\subsection{Subsection}
-\subsubsection{Subsubsection}
-
-Bulleted lists look like this:
-\begin{itemize}
-\item	First bullet;
-\item	Second bullet;
-\item	Third bullet.
-\end{itemize}
-
-Numbered lists can be added as follows:
-\begin{enumerate}
-\item	First item; 
-\item	Second item;
-\item	Third item.
-\end{enumerate}
-
-The text continues here.
-
-\subsection{Figures, Tables and Schemes}
-
-All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, etc.
-
-%\begin{figure}[H]
-%\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
-%\caption{This is a figure.
-Schemes follow the same formatting.
-If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
-(\textbf{b}) Description of what is contained in the second panel.
-Figures should be placed in the main text near to the first time they are cited.
-A caption on a single line should be centered.\label{fig1}}
-%\end{figure}
-
-% The MDPI table float is called specialtable
-\begin{specialtable}[H] 
-\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
-%%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
-\begin{tabular}{ccc}
-\toprule
-\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-\midrule
-Entry 1		& Data			& Data\\
-Entry 2		& Data			& Data\\
-\bottomrule
-\end{tabular}
-\end{specialtable}
-
-%\begin{listing}[H]
-%\caption{Title of the listing}
-%\rule{\columnwidth}{1pt}
-%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
-%\rule{\columnwidth}{1pt}
-%\end{listing}
-
-Text.
-
-Text.
-
-\subsection{Formatting of Mathematical Components}
-
-This is the example 1 of equation:
-\begin{equation}
-a = 1,
-\end{equation}
-the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
-%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
-This ensures correct linenumbering.
-The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
-
-This is the example 2 of equation:
-\end{paracol}
-\nointerlineskip
-\begin{equation}
-a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
-\end{equation}
-
-% Example of a figure that spans the whole page width (the commands \widefigure and \begin{paracol}{2}, \linenumbers, and\switchcolumn need to be present). The same concept works for tables, too.
-%\begin{figure}[H]	
-%\widefigure
-%\includegraphics[width=15 cm]{Definitions/logo-mdpi}
-%\caption{This is a wide figure.\label{fig2}}
-%\end{figure}  
-\begin{paracol}{2}
-\linenumbers
-\switchcolumn
-
-Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
-%% Example of a theorem:
-\begin{Theorem}
-Example text of a theorem.
-\end{Theorem}
-
-The text continues here. Proofs must be formatted as follows:
-
-%% Example of a proof:
-\begin{proof}[Proof of Theorem 1]
-Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
-\end{proof}
-The text continues here.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses.
-The findings and their implications should be discussed in the broadest context possible.
-Future research directions may also be highlighted.
 

### 2021-07-28 2021/Entropy/entropy.tex
Before translation of Sections 1-3

@@
 % Full title of the paper (Capitalized)
-\Title{Title}
+\Title{Ускорение метода глобальной оптимизации за счет выделения локальных экстремумов на основе машинного обучения}
+%		   Acceleration of Global Optimization Algorithm by Detecting Local Extrema Based on Machine Learning
 
 % MDPI internal command: Title for citation in the left column
-\TitleCitation{Title}
+\TitleCitation{}
 
@@
 % Authors, for the paper (add full first names)
-\Author{Konstantin Barkalov $^{1}$*\orcidA{}, Ilya Lebedev $^{1}$\orcidB{} and Victor Gergel $^{1}$\orcidC{}}
+\Author{Konstantin Barkalov $^{1}$*\orcidA{} and Ilya Lebedev $^{1}$\orcidB{}}
 
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Konstantin Barkalov, Ilya Lebedev and Victor Gergel}
+\AuthorNames{Konstantin Barkalov and Ilya Lebedev}
 
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Barkalov, K.; Lebedev, I.; Gergel, V.}
+\AuthorCitation{Barkalov, K.; Lebedev, I.}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
@@
 \address[1]{%
-$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhni Novgorod, Russia; ilya.lebedev@itmm.unn.ru (I.L.); gergel@unn.ru (V.G.)}
+$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhni Novgorod, Russia; ilya.lebedev@itmm.unn.ru (I.L.)}
 %$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
@@
 % Contact information of the corresponding author
-\corres{Correspondence: konstantin.barkalov@itmm.unn.ru}
+\corres{Correspondence: konstantin.barkalov@itmm.unn.ru (K.B.)}
 %; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}
@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{A single paragraph of about 200 words maximum.
-For research articles, abstracts should give a pertinent overview of the work.
-We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusion: indicate the main conclusions or interpretations.
-The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}
+\abstract{
+Рассматриваются задачи глобальной оптимизации и численные методы их решения.
+Задачи указанного класса являются вычислительно-трудоемкими, т.к.
+целевая функция может являться многоэкстремальной, недифференцируемой, и, как правильно, заданной в виде «черного ящика».
+В проведенном исследовании использовался детерминированный алгоритм поиска глобального экстремума, который не основан на идях мультистарта и nature-inspired алгоритмов.
+В статье приведены вычислительные правила одномерного алгоритма и схема вложенной оптимизации, которая позволяет применять его для решения сложных многомерных задач.
+Отметим, что сложность решения задач глобальной оптимизации существенным образом зависит от наличия многих локальных экстремумов.
+В работе предложен алгоритм выделения областей притяжения локальных минимумов, основанный на использовании методов машинного обучения (в частности, деревьев решений).
+Использование в выделенных областях алгоритмов локальной оптимизации позволяет значительно ускорить сходимость глобального поиска за счет снижения числа поисковых испытании в окрестностях локальных минимумов.
+Приведены результаты вычислительных экспериментов, проведенных на нескольких сотнях задач глобальной оптимизации разной размерности, которые подтверждают эффект ускорения сходимости (в терминах числа поисковых испытаний, требующихся для решения задачи с заданной точностью).
+}
 
 % Keywords
-\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}
+\keyword{global optimization; local optimization; multiextremal problems; numerical methods; approximation; decision trees}
 
@@
 Успешное применение методов машинного обучения (ML) для решения широкого спектра задач приводит к созданию новых подходов к эффективному использованию идей ML в различных областях приложений. 
-
 Примером класса задач, в которых методы машинного обучения продемонстрировали свою эффективность, являются задачи выявления основных свойств исследуемых явлений (например, физических, экономических или социальных), которые характеризуются стохастической природой или наличием скрытых параметров \cite{Golovenkin2020,Gonoskov2019}.
-
 Методы машинного обучения успешно используются и для решения сложных задач вычислительной математики, например, для simulation of dynamical systems \cite{Seleznev2019}, решения ordinary, partial or stochastic differential equations \cite{Lagaris1998,Blechschmidt2021,Xu2020} .
@@
 Можно выделить несколько подходов к построению численных методов решения задач глобальной оптимизации. 
-
-%Алгоритмы на основе метамоделей
-
+%Надо ли добавить алгоритмы на основе метамоделей ?
 Ряд алгоритмов основан на идее мультистарта: запуск локального поиска либо начиная с разных стартовых точек, либо с использованием разных параметров.
@@
 
-Будем рассматривать задачи глобальной оптимизации вида
+В данной работе мы будем рассматривать задачи глобальной оптимизации вида
 \begin{eqnarray}\label{main_problem}
@@
 \end{eqnarray}
-Задача \ref{main_problem} рассматривается в предположении, что целевая функция является многоэкстремальной, задана как ``черный ящик'', а вычисление ее значений связано с решением задачи численного моделирования и является трудоемкой операцией.
+Задача (\ref{main_problem}) рассматривается в предположении, что целевая функция является многоэкстремальной, задана как ``черный ящик'', а вычисление ее значений связано с решением задачи численного моделирования и является трудоемкой операцией.
 
@@
 Для преодоления сложностей, вызванных размерностью решаемой задачи, в многоэкстремальной оптимизации широко используются различные подходы к уменьшению размерности.  
-Например, симплексное или диагональное разбиение области поиска позволяет использовать для решения исходной многомерной задачи методы решения одномерных задач (see, for example, \cite{PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
+Например, simplicial or diagonal partition области поиска позволяет использовать для решения исходной многомерной задачи методы решения одномерных задач (see, for example, \cite{PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
 Известным подходом к редукции размерости является также использование Peano space-filling curves mapping the multidimensional domain onto one-dimensional interval \cite{Strongin2000}.
@@
 
-Приведем описание алгоритма глобального поиска (АГП) для решения базовой задачи в соответствии с \cite{Strongin2000}.
-В процессе своей работы АГП порождает последовательность точек $x^i$, в которых вычисляются значения целевой функции $z^i=\varphi(x^i)$. 
-Будем называть процесс вычисления значения целевой функции \textit{испытанием}.
+Приведем описание global search algorithm (GSA) для решения базовой задачи в соответствии с \cite{Strongin2000}.
+В процессе своей работы GSA порождает последовательность точек $x^i$, в которых вычисляются значения целевой функции $z^i=\varphi(x^i)$. 
+Будем называть процесс вычисления значения целевой функции \textit{trial}.
 
@@
 Теоретические условия, определяющие сходимость алгоритма, представлены в \cite{Strongin2000}. 
+Работа алгоритма при минимизации конкретной многоэкстремальной функции, которая задаются в соответствии с формулой (\ref{hill}), проиллюстрирована на fugure \ref{fig1}.
+При запуске алгоритма использовался параметр $r=2.2$ из (\ref{R}) и значение $\epsilon = 10^{-3}$ в критерии остановки метода.
+На рисунке \ref{fig1} изображен график objective function и точки 71 поискового испытания, потребовавшихся GSA для решения задачи с указанной точностью.
+Наглядно видна проблема всех методов глобальной оптимизации -- концентрация точек в окрестностях локальных минимумов задачи, которые не являются глобальным решением.
 
-%Пример работы алгоритма - функция Хилла. Эту же задачу надо будет показать и в след.разделе
+\begin{figure}[H]
+\includegraphics[width=1.0\linewidth]{HillAGP90.png}
+\caption{Минимизация функции вида (\ref{hill}) с помощью GSA.}
+\label{fig1}
+\end{figure}
 
@@
 Далее сравниваются значения $c_j$, соответствующие соседним подобластям. Если выполняется одно из условий
-\begin{gather*}\label{LocalPointC}
-c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3} \leq c_{j+4}, \; j=1, \\ 
-c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3}, \; j=2,  \\ 
+\begin{gather}\label{LocalPointC}
+c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3} \leq c_{j+4}, \; j=1; \nonumber \\ 
+c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3}, \; j=2;  \nonumber \\  
 c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2}, \; 3 \leq j \leq J-2; \\ 
-c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1}, \; j=J-1, \\ 
-c_{j-4} \geq c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j}, \; j=J, 
-\end{gather*}
+c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1}, \; j=J-1; \nonumber \\ 
+c_{j-4} \geq c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j}, \; j=J; \nonumber 
+\end{gather}
 то область $D_j$ считается областью притяжения локального минимума. В ней можно запускать локальный поиск, а впоследствии -- исключить данную подобласть из глобального поиска.
@@
 
-Опишем теперь модифицированный алгоритм глобального поиска (GSA), использующий деревья решений для выделения и исключения областей притяжения локальных минимумов; будем в дальнейшем ссылаться на данный алгоритм как на GSA-DT.
+Опишем теперь модифицированный алгоритм глобального поиска, использующий деревья решений для выделения и исключения областей притяжения локальных минимумов; будем в дальнейшем ссылаться на данный алгоритм как на GSA-DT.
 
@@
 
-Рисунки \ref{fig1} и figure \ref{fig2} иллюстрируют работу алгоритма при минимизации конкретной многоэкстремальной функции, которая задаются в соответствии с формулой (\ref{hill}).
-При запуске алгоритма использовался параметр $r=2.2$ из (\ref{R}) и значение $\epsilon = 10^{-3}$ в критерии остановки метода.
-
-На рисунке \ref{fig1} изображен график objective function и точки 71 поискового испытания, потребовавшихся GSA для решения задачи с указанной точностью.
-Наглядно видна концентрация точек в окрестностях локальных минимумов задачи.
-Рисунок \ref{fig2} соответствуют работе алгоритма GSA-DT.
+Для примера рассмотрим работу алгоритма GSA-DT при минимизации той же многоэкстремальной функции, что представлена на fugire \ref{fig1}.
+При запуске алгоритма использовались такре же параметры: параметр $r=2.2$ из (\ref{R}) и значение $\epsilon = 10^{-3}$ в критерии остановки метода.
+Рисунок \ref{fig2} иллюстрирует работу алгоритма GSA-DT.
 Кроме графика целевой функици на нем изображена кусочно-постоянная аппроксимация вида (\ref{dtree}), построенная на финальной стадии поиска.
@@
 
-%далее нужны иллюстрации
 \begin{figure}[H]
-\label{fig1}
-\includegraphics[width=1.0\linewidth]{HillAGP90.png}
-\caption{Минимизация функции вида (\ref{hill}) с помощью GSA.}
-\end{figure}
-
-\begin{figure}[H]
-\label{fig2}
 \includegraphics[width=1.0\linewidth]{HillTree90.png}
 \caption{Минимизация функции вида (\ref{hill}) с помощью GSA-DT.}
+\label{fig2}
 \end{figure}
@@
 
-Следующую серию экспериментов проведем при решении многомерных задач. 
+В следующей серии экспериментов решались многомерные задачи. 
 Известным генератором тестовых задач многоэкстремальной оптимизации является генератор GLKS \cite{Gaviano2003}.
@@
 \begin{specialtable}[H] 
-	\caption{Average number of iterations GKLS problem}\label{table:average_GKLS}
+	\caption{Решение задач GKLS с точностью $\epsilon = 10^{-2}$}\label{table:average_GKLS01}
 	\center
-\begin{tabular}{ccccccc}
-\hline\noalign{\smallskip}
+\begin{tabular}{cccc}
+\toprule
+        & $N=2$ & $N=3$  & $N=4$    \\
+\midrule
+GSA     &  937  &  12716 &  206869  \\
+GSA-DT  &  653  &  9204  &  156190  \\
+\bottomrule
+\end{tabular}
+\end{specialtable}
+
+\begin{specialtable}[H] 
+	\caption{Решение задач GKLS с точностью $\epsilon = 2 \cdot 10^{-3}$}\label{table:average_GKLS002}
+	\center
+\begin{tabular}{cccc}
 \toprule
 
-                      & \multicolumn{2}{l}{2D}                                                          & \multicolumn{2}{l}{3D}                                                          & \multicolumn{2}{l}{4D}                                  \\
-                      & \textbf{0,002} & \textbf{0,01} & \textbf{0,002} & \textbf{0,01} & \textbf{0,002} & \textbf{0,01} \\
+       & $N=2$ & $N=3$  & $N=4$    \\
 \midrule
-DiRect                &   39,84(0)  & 29,92
- (6)                            &                                        &                                       &                &                                       \\
-Базовый АГП           & 1820                                   & 841                                   & 53251,32                               & 8905,03                               & 665013,15      & 135972,78 (99)                        \\
-Адаптивная схема      & 1489,01                                & 937,20                                & 69764,39                               & 12716,21                              & 583903,06      & 206869,83                             \\
-АГП с деревом решений & 831,77                                 & 653,94                                & 10776,87                               & 9204,16                               & 173155,3       & 156190,43                              \\
+GSA    & 1489  & 69764  & 583903  \\
+GSA-DT & 831   & 10776  & 173155   \\
 \bottomrule

### 2021-08-03 2021/ICNAAM/2021_ICNAAM.tex
Corrections in Russian

@@
 \begin{abstract}
-The paper presents the results of solving the problem of simulation and research of processes taking place in a fuel system injector.
-Simulation has been carried using the tools of the OpenFOAM system.
+The paper presents results of solving the problem of simulation and research of processes taking place in a fuel system injector.
+Simulation has been carried using tools of the OpenFOAM system.
 The problem of differential pressure minimization at the injector inlet and outlet was set and solved.
-This problem was a Lipschitz optimization problem with a black-box type objective function.
+This problem was considered as a Lipschitz optimization problem with a black-box type objective function.
 To solve it, we used an efficient global search algorithm implemented in the Globalizer system.
@@
 For this purpose, CAD/CAM/CAE-systems are widely used.
-A well-known example of open source software of this kind is OpenFOAM, an open platform for numerical simulation of continuum mechanics problems \cite{OpenFOAM}).
+A well-known example of open source software of this kind is OpenFOAM, an open platform for numerical simulation of continuum mechanics problems \cite{OpenFOAM}.
 
@@
 The optimal parameter selection problems under consideration are characterized by the fact that the objective function $\varphi(y)$ is not defined analytically; there is only some algorithm to calculate its values at the points of the domain $D$.
-In this case, one search trial (calculating the function value in a particular point of the search domain) is a computationally intensive operation \cite{Kalyulin2017}.
+In this case, one search trial (calculating the function value in a particular point of the search domain) is a computationally intensive operation \cite{Sergeyev1999,Kvasov2013,Kalyulin2017,Sergeyev2018,Paulavicius2020,Lera2021}.
 
@@
 
-A typical assumption used in many global optimization methods \cite{Evtushenko2013,Jones2009,Zilinskas2010,Pinter1996} is the assumption that the objective function $\varphi(y)$ satisfies the Lipschitz condition
+A typical assumption used in many global optimization methods \cite{Sergeyev2013,Evtushenko2013,Jones2009,Zilinskas2010,Pinter1996,Grishagin2016_1,Sergeyev2021} is the assumption that the objective function $\varphi(y)$ satisfies the Lipschitz condition
 \[
@@
 \]
+% Предположение такого рода является достаточно естественным для многих прикладных задач, поскольку относительные вариации целевой функции, характеризующей моделируемую систему, обычно не могут превышать некоторый порог, определяемый ограниченной энергией изменений в системе.
 
@@
 \]
+As follows from the properties of the space-filling curve \cite{Sergeyev2013}, the reduced one-dimensional function $\varphi(y(x))$ satisfies a uniform H{\"o}lder condition
+\[
+\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}, \; x_1,x_2 \in [0,1],
+\]
+where the H{\"o}lder constant $H$ is related to the Lipschitz constant $L$ by the relationship $H=2L\sqrt{N+3}$.
 
@@
 Here, we present a scheme of the basic global search algorithm (GSA).
-This is a Divide-the-Best algorithm, i.e.
+This is a Divide-the-Best algorithm \cite{Sergeyev1998}, i.e.
 it is aimed to conduct a subsequent search trial in the most promising (from the point of view of global minimum detection) search subdomain.
@@
 
-Step 2. For each interval $(x_{i-1},x_i), \; 1\leq i \leq k,$ compute the characteristics $R(i)$ using some formulae.
+Step 2. For each interval $(x_{i-1},x_i), \; 1\leq i \leq k+1,$ compute the characteristics $R(i)$ using some formulae.
 
-Step 3. Find the interval $(x_{t-1},x_t),$ with the maximum characteristic 
+Step 3. Find the interval $(x_{t-1},x_t),$ with the maximum characteristic \cite{Grishagin1997}
 \[
@@
 Step 5. Check termination condition  $\left(x_t-x_{t-1}\right)^{1/N}<\epsilon$, where $t$ is the number of the interval with the maximum characteristic and $\epsilon>0$ is the predefined accuracy.
+
+The global search algorithm has the folloing interpretation.
+Using trial points it is possible to design a minorant for which characteristic $R(i)$ is the minimum value of the minorant in the interval $(x_{i-1}, x_i ), 1 \leq i \leq k + 1,$ taken with the opposite sign.
+Then new trial is carried out in the interval where the global minimizer of the current minorant is situated and the new trial point $x^{k+1}$ coincides with it.
 
@@
 \begin{enumerate}
-\item Mathematical formulation.
+\item \textbf{Mathematical formulation}.
 This includes a procedure for selecting equations describing the dynamics of a fluid or gas, setting boundary conditions and initial conditions.
 It does not change from iteration to iteration.
-\item Parameterization. Selection of varying parameters that will vary from iteration to iteration.
-\item Generating the geometry and the computational grid. In line with Parameterization, a procedure for renewing the geometry and computational grid is performed.
-\item Calculation of hydrodynamic fields. At each iteration, the system of equations describing fluid dynamics is solved numerically.
-\item Interpretation. Physical interpretation of the results after the optimization procedure.
+\item \textbf{Parameterization}. Selection of varying parameters that will vary from iteration to iteration.
+\item \textbf{Generating the geometry and the computational grid}. In line with Parameterization, a procedure for renewing the geometry and computational grid is performed.
+\item \textbf{Calculation of hydrodynamic fields}. At each iteration, the system of equations describing fluid dynamics is solved numerically.
+\item \textbf{Interpretation}. Physical interpretation of the results after the optimization procedure.
 \end{enumerate}
@@
 The geometry is parametrized by increasing the distance between the red points symmetrically with respect to the channel centreline.
+%Таким образом, в задаче имеется один варьируемый параметр (диаметр сечения канала в его средней части), значения которого изменяются в диапазоне $[0,10]$.
 
@@
 The process of finding the effective channel thickness took 3395 seconds on an Intel(R) Xeon(R) CPU X5670 2.93 GHz, during which time 69 iterations of the global search algorithm were performed.
+% Для решения задачи был использован смешанный локально-глобальный алгоритм \cite{Barkalov2010}, реализованный в Globalizer.
+% При этом в критерии остановки метода было использовано значение $\epsilon = 10^{-3}$.
 The horizontal component of the velocity at the entrance to the computational domain $\vec{U_i}$  was 10 m/s.
@@
 As a result of solving the problem, the optimum channel thickness (0.59 cm) was found, giving the minimum fuel consumption.
+
+%Найденное решение является удовлетворительным с точки зрения решаемой задачи.
+Например, другие решения, которые получаются при запуске методов локальной оптимизации для решения данной задачи, зависят от выбора стартовой точки.
+И значение целевой функции при этом получается не меньше, чем значение в точке, найденной Globalizer.
 

### 2021-08-03 2021/Entropy/entropy.tex
two new figures

@@
 
+Для примера рассмотрим работу алгоритмов GSA и GSA-DT при решении одной из задач, порожденных генератором GKLS. 
+При запуске алгоритмов использовались одинаковые параметры: $r=3.0$ из (\ref{R}) и $\epsilon = 10^{-2}$ в критерии остановки метода.
+
+Линии уровня оптимизируемой функции, изображенные на рисунках, наглядно показывают многоэкстремальность функции.
+
+На figure \ref{fig3} изображены точки 247 испытаний, потребовавшихся алгоритму для решения задачи с указанной точностью.
+Точкой красного цвета отмечено точное решение задачи, а точкой желтого цвета - лучшее его приближение, найденное алгоритмом.
+
+При запуске алгоритма использовались такие же параметры: параметр $r=2.2$ из (\ref{R}) и значение $\epsilon = 10^{-3}$ в критерии остановки метода.
+Рисунок \ref{fig2} иллюстрирует работу алгоритма GSA-DT.
+Кроме графика целевой функици на нем изображена кусочно-постоянная аппроксимация вида (\ref{dtree}), построенная на финальной стадии поиска.
+Черные точки на графике соответствуют фазе глобального поиска, зеленые точки соответствуют работе локального метода.
+Всего методу GSA-DT потребовалось 49 испытаний для решения задачи; накопление точек испытаний в окрестностях локальных минимумов отсутствует.
+
+\begin{figure}[H]
+\includegraphics[width=0.6\linewidth]{GKLSAdaptiv6_30line.png}
+\caption{Решение задачи GKLS с помощью алгоритма GSA.}
+\label{fig3}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=0.6\linewidth]{GKLSTree6_30line.png}
+\caption{Решение задачи GKLS с помощью алгоритма GSA-DT.}
+\label{fig3}
+\end{figure}
+
 \begin{specialtable}[H] 

### 2021-08-04 2021/Entropy/entropy.tex
main part of the text in Russian is ready

--- 2021-07-28 2021/Entropy/entropy.tex
@@
 \abstract{
-Рассматриваются задачи глобальной оптимизации и численные методы их решения.
+В статье рассматриваются задачи глобальной оптимизации и численные методы их решения.
 Задачи указанного класса являются вычислительно-трудоемкими, т.к.
 целевая функция может являться многоэкстремальной, недифференцируемой, и, как правильно, заданной в виде «черного ящика».
-В проведенном исследовании использовался детерминированный алгоритм поиска глобального экстремума, который не основан на идях мультистарта и nature-inspired алгоритмов.
-В статье приведены вычислительные правила одномерного алгоритма и схема вложенной оптимизации, которая позволяет применять его для решения сложных многомерных задач.
+В проведенном исследовании использовался детерминированный алгоритм поиска глобального экстремума, который не основан на идеях мультистарта и nature-inspired algorithms.
+В статье приведены вычислительные правила одномерного алгоритма и nested optimization scheme, которая позволяет применять его для решения сложных многомерных задач.
 Отметим, что сложность решения задач глобальной оптимизации существенным образом зависит от наличия многих локальных экстремумов.
@@
 \]
-%Отметим, что значение константы $L$ обычно является априори неизвестным, что делает ее оценку одной из ключевых проблем при построении методов липшицевой оптимизации.$\overline{\underline{\text{XX}}}$
 Предположение липшицевости целевой функции типично для многих подходов к разработке оптимизационных алгоритмов \cite{Jones1993,Pinter1996,Zilinskas2008,Evtushenko2009}.
@@
 Как результат, при минимизации существенно многоэкстремальных функций численный метод должен построить покрытие области поиска и число узлов этого покрытия увеличивается экспоненциально с ростом размерности.
-Эта особенность обуславливает высокую трудоемкость решения задач многоэкстремальной оптимимизации, и размерность является решающим фактором, влияющим на сложность их решения.
+Эта особенность обуславливает высокую трудоемкость решения задач многоэкстремальной оптимизации, и размерность является решающим фактором, влияющим на сложность их решения.
 
@@
 The nested optimization scheme, с одной стороны, не ухудшает свойств минимизируемой функции (в отличие от редукции с использованием Peano curves), а с другой стороны, не требует использования сложных структур данных для поддержки симплексных или диагональных разбиений допустимой области.
-Одновременно с этим схема вложенной оптимизации позволяет свести решение многомерной задачи к решению набора стандартных одномерных задач и использовать широкий спектр алгоритмов одномерной глобальной оптимизации для их решения \cite{}.
+Одновременно с этим схема вложенной оптимизации позволяет свести решение многомерной задачи к решению набора стандартных одномерных задач и использовать широкий спектр алгоритмов одномерной глобальной оптимизации для их решения.
 
@@
 Теоретические условия, определяющие сходимость алгоритма, представлены в \cite{Strongin2000}. 
-Работа алгоритма при минимизации конкретной многоэкстремальной функции, которая задаются в соответствии с формулой (\ref{hill}), проиллюстрирована на fugure \ref{fig1}.
+Работа алгоритма при минимизации конкретной многоэкстремальной функции, которая задаются в соответствии с формулой (\ref{hill}), проиллюстрирована на figure \ref{fig1}.
 При запуске алгоритма использовался параметр $r=2.2$ из (\ref{R}) и значение $\epsilon = 10^{-3}$ в критерии остановки метода.
@@
 Построить регрессию можно также и с использованием такого мощного инструмента, как искусственные нейронные сети. 
-Для построения регрессии могут быть использованы сети разного типа, например, многослойный персептрон, сеть радиальных базисных функций, и т.д. 
+Для построения регрессии могут быть использованы сети разного типа, например, multilayer perceptron, radial basis function network, и т.д. 
 Однако в обоих этих случаях модель (сплайн или же нейронная сеть) сама становится сложной для того анализа, который требуется выполнить в решаемой проблеме (выделение областей притяжения локальных экстремумов).
@@
 Рисунок \ref{fig2} иллюстрирует работу алгоритма GSA-DT.
-Кроме графика целевой функици на нем изображена кусочно-постоянная аппроксимация вида (\ref{dtree}), построенная на финальной стадии поиска.
+Кроме графика целевой функции на нем изображена кусочно-постоянная аппроксимация вида (\ref{dtree}), построенная на финальной стадии поиска.
 Черные точки на графике соответствуют фазе глобального поиска, зеленые точки соответствуют работе локального метода.
@@
 
-Для формального описания многошаговой схемы введем семейство функций, определяемых в соответствии с соотношениями 
+Для формального описания nested optimization scheme введем семейство функций, определяемых в соответствии с соотношениями 
 \begin{equation}\label{nested_N}
@@
 
-Адаптивная многошаговая схема редукции размерности изменяет порядок решения подзадач: они будут решаться не по одной (в соответствии с их иерархией в дереве задач), а одновременно, т.е.
+Adaptive nested optimization scheme  редукции размерности изменяет порядок решения подзадач: они будут решаться не по одной (в соответствии с их иерархией в дереве задач), а одновременно, т.е.
 будет существовать некоторое множество подзадач, находящихся в процессе решения.
@@
 В частности, если использовать точность $\epsilon = 10^{-2}$, то метод DIRECT останавливается слишком рано, и не находит глобального решения во многих задачах.
+Поэтому в дальнейших экспериментах, в которых решаются многомерные задачи, мы не будем использовать DIRECT для сравнения, т.к.
+при решении многомерных задач с остановкой по точности данный метод обеспечиват корректное решение не более чем $50\%$ задач.
 
@@
 Известным генератором тестовых задач многоэкстремальной оптимизации является генератор GLKS \cite{Gaviano2003}.
-С его помощью можно порождать задачи оптимизации с заданными свойствами: числом локальных экстремумов, их областями притяжения, точкой глобального минимума и значением целевой функции в ней, и т.д.
+С его помощью можно порождать тестовые функции с заданными свойствами: числом локальных экстремумов, их областями притяжения, точкой глобального минимума и значением целевой функции в ней, и т.д.
+Процедура генерирования тестовых функций состоит в переопределении выпуклой квадратичной функции (параболоида) при помощи полиномов. Тестовые функции определяются пятью параметрами:
+\begin{itemize}
+	\item размерность задачи $N$;
+	\item число локальных минимумов $l$;
+	\item значение глобального минимума $f^*$;
+	\item радиус области притяжения global optimizer $\rho^*$;
+	\item расстояние между global optimizer и вершиной параболоида $d^*$.
+\end{itemize}
+Изменяя указанные параметры, можно создавать тестовые классы с различными свойствами.
+Например, при фиксированных размерности задачи и числи локальных минимумов более сложный класс может быть сгенерирован путем сужения области притяжения точки глобального минимума или путем увеличения расстояния между этой точкой и вершиной параболоида.
+В проведенных экспериментах использовались значения $l=10$, $f^*=-1$, $\rho^*=0.2$ и $d^*=0.9$.
+
+Для примера рассмотрим работу алгоритмов GSA и GSA-DT при решении одной из двумерных задач, порожденных генератором GKLS. 
+Линии уровня оптимизируемой функции, изображенные на figures \ref{fig3} и \ref{fig4}, наглядно показывают наличие десяти локальных экстремумов.
+При запуске алгоритмов использовались одинаковые параметры: $r=3.0$ из (\ref{R}) и $\epsilon = 10^{-2}$ в критерии остановки метода.
+На figures \ref{fig3} и \ref{fig4} точками черного цвета изображены точки search trials, выполненных методами в процессе решения задачи.
+При этом алгоритму GSA потребовалось 247 испытаний, тогда как алгоритму GSA-DT хватило 138 испытаний.
+Точкой красного цвета на рисунках отмечено точное решение задачи, а точкой желтого цвета - лучшее его приближение, найденное алгоритмом. 
+На рисунке \ref{fig4} точками зеленого цвета отмечены испытания, выполненные в рамках локального поиска. 
+Наглядно видно, что в методе с использованием деревьев решений для выделения областей притяжения локальных минимумов снимается проблема накопления точек испытаний в районе локальных экстремумов, присущая исходному алгоритму глобального поиска.
+
+\begin{figure}[H]
+\includegraphics[width=0.6\linewidth]{GKLSAdaptiv6_30line.png}
+\caption{Решение задачи GKLS с помощью алгоритма GSA.}
+\label{fig3}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=0.6\linewidth]{GKLSTree6_30line.png}
+\caption{Решение задачи GKLS с помощью алгоритма GSA-DT.}
+\label{fig4}
+\end{figure}
+
+С помощью генератора GKLS было порождено по 300 тестовых задач размерностей $N=2,3,4$ (по 100 задач каждой размерности).
+Полученные серии задач были решены с помощью алгоритмов GSA и GSA-DT с параметром $r=5.0$ из (\ref{R}).
+Указанное значение параметра $r$ обеспечивает решение $100\%$ задач, при меньших значениях параметра некоторые задачи не были решены корректно.
+Отметим, что метод DIRECT
+
+В таблицах \ref{table:average_GKLS01} и \ref{table:average_GKLS002} приведено среднее число trials, потребовавшихся методам GSA и GSA-DT для корректного решения всех задач с точностью $\epsilon = 10^{-2}$ и $\epsilon = 2 \cdot 10^{-3}$ соответственно.
+
+Данные из таблиц подтверждают, что алгоритм глобального поиска, использующий выделение локальных экстремумов на основе машинного обучения, обеспечивает более быстрое решение многоэкстремальных задач, чем базовый алгоритм глобального поиска.
+При решении задач с грубой точностью ускорение составляет порядка $30\%$, тогда как при решении задач с высокой точностью процесс решения ускоряется от 2 до 6 раз.
 
@@
 
-В таблицах \ref{table:average_GKLS} приведено среднее число итераций полученное на  задачах GKLS, для алгоритмов Direct, АГП, АГП с адаптивной схемой и АГП с адаптивной схемой с использованием локального метода запущенного по результатам работы дерева решений.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
+В статье рассмотрен эффективный детерминированный метод решения задач многоэкстремальной оптимизации -- the information-statistical global search algorithm. 
+Для данного алгоритма предложен новый способ ускорения его работы (в терминах числа итераций, требующихся для решения задачи с заданной точностью).
+Указанный способ основан на выявлении областей притяжения локальных минимумов целевой функции с использованием методов машинного обучения.
+Выявление областей притяжения и запуск в этих областях локального поиска позволяет существенным образом сократить число испытаний, требующееся методу для достижения глобальной сходимости.
+В рамках исследуемого подхода решение многомерных задач сводится к решению серии информационно-связанных одномерных подзадач, поэтому ключевым моментом является выявление локальных минимумов в одномерных задачах.
+Для этого используется аппроксимация целевой функции по точкам поисковых испытаний, построенная с помощью деревьев решений.
+Проведены вычислительные эксперименты на серии тестовых задач разной размерности с целью сравнения скорости работы исходного global search algorithm (GSA) и его модификации, которая использует деревья решений для выделения локальных минимумов целевой функции (GSA-DT).
+Результаты экспериментов показывают, что использование алгоритма GSA-DT может значительно (до 6 раз) сократить число испытаний, необходимое для решения задачи с заданной точностью.
+
+Направлением дальнейших исследований будет использование более сложных моделей целевой фукнции для получения более точной ее аппроксимации.
+В качестве такого аппроксиматора планируется использовать artificial neural networks.
+Это потребует разработки новых методов выделения локальных экстремумов, т.к.
+аппроксимация функции с помощью нейронной сети является более сложной с этой точки зрения.
+Также планируется уделить внимание вопросам достоверности результатов, полученных с помощью методов машинного обучения.
+Для решенных модельных задач использование методов машинного обучения показывает хорошие результаты, однако вопрос о том, сохранится ли этот эффект в более сложных задачах, остается открытым.
 

### 2021-08-06 2021/Entropy/entropy.tex
before translation

@@
 \Title{Ускорение метода глобальной оптимизации за счет выделения локальных экстремумов на основе машинного обучения}
-%		   Acceleration of Global Optimization Algorithm by Detecting Local Extrema Based on Machine Learning
+		   %Acceleration of Global Optimization Algorithm by Detecting Local Extrema Based on Machine Learning
 
@@
 \newcommand{\orcidauthorB}{0000-0002-8736-0652} % Add \orcidB{} behind the author's name
-\newcommand{\orcidauthorC}{0000-0002-4013-2329} % Add \orcidB{} behind the author's name
+%\newcommand{\orcidauthorC}{0000-0002-4013-2329} % Add \orcidB{} behind the author's name
 
@@
 В работе предложен алгоритм выделения областей притяжения локальных минимумов, основанный на использовании методов машинного обучения (в частности, деревьев решений).
-Использование в выделенных областях алгоритмов локальной оптимизации позволяет значительно ускорить сходимость глобального поиска за счет снижения числа поисковых испытании в окрестностях локальных минимумов.
-Приведены результаты вычислительных экспериментов, проведенных на нескольких сотнях задач глобальной оптимизации разной размерности, которые подтверждают эффект ускорения сходимости (в терминах числа поисковых испытаний, требующихся для решения задачи с заданной точностью).
+Использование в выделенных областях алгоритмов локальной оптимизации позволяет значительно ускорить сходимость глобального поиска за счет снижения числа search trials в окрестностях локальных минимумов.
+Приведены результаты вычислительных экспериментов, проведенных на нескольких сотнях задач глобальной оптимизации разной размерности, которые подтверждают эффект ускорения сходимости (в терминах числа search trials, требующихся для решения задачи с заданной точностью).
 }
@@
 
-Numerical experiments were performed on the Lobachevsky supercomputer of the University of Nizhni Novgorod (operating system -- CentOS 6.4, management system -- SLURM).
+Numerical experiments were performed on the Lobachevsky supercomputer of the University of Nizhni Novgorod (operating system CentOS 7.2, management system SLURM).
 One supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM.
@@
 a total of 16 CPU cores are available on the node).
+Все используемые алгоритмы были реализованы на C++, для сборки на суперкомпьютере использовался компилятор GCC 5.5.0.
 
@@
 При этом подразумевается, что очередная решаемая задача генерируется посредством некоторого алгоритма.
-Типичными примерами таких классов тестовых функций являются функции  Шекеля и Хилла. 
+Типичными примерами таких классов тестовых функций являются функции Шекеля и Хилла (Shekel и Hill). 
 Первый из них (обозначит его $F_{SH}$) основан на формуле 
@@
 \end{equation}
-где значения параметров  $A_j,\: B_j,\: 1 \le j \le 14$, независимо и равномерно распределены в интервале $[-1 ,1 ]$.
+где значения параметров  $A_j,\: B_j,\: 1 \le j \le 14$, независимо и равномерно распределены в интервале $[-1,1]$.
 
 Проведем сравнение базового global search algorithm (GSA) и его модификации, использующей деревья решений (GSA-DT), c известным алгоритмом глобальной оптимизации DIRECT \cite{Jones2009}.  
-Сравнение методов будем проводить при решении 1000 задач из классов $F_{SH}$ и $F_{HL}$.
+Для построения деревьев решений в алгоритме GSA-DT была использована библиотека OpenCV 4.5.1 (класс cv::ml::DTrees).
+Использовалось одиночное дерево решения для задачи построения функции регрессии.
+В результате получалась кусочно-постоянная аппроксимация целевой функции, в которой каждому листу дерева соответствовала одна или несколько trial points.
+Для построения дерева максимальная глубина (MaxDepth) не ограничивалась, точность построения дерева (RegressionAccuracy) составляла $10^{-3}$, минимальное количество точек испытания в узле дерева (MinSampleCount) равнялось одному, все остальные параметры были заданы по умолчанию.
+Остановка построения дерева происодила, if all absolute differences between an estimated value in a node and values of train samples in this node are less than accuracy.
 
+Сравнение методов глобальной оптимизации проводилось при решении 100 задач из классов $F_{SH}$ и $F_{HL}$.
 Задачу будем считать решенной корректно в случае, если после остановки метода по точности (т.е.
@@
 
-В таблицах \ref{table:average_Shekel} и \ref{table:average_Hill} приведено число поисковых испытаний, которые в среднем требовалось для минимизации функций Шекеля и Хилла с разной точностью поиска.
+В таблицах \ref{table:average_Shekel} и \ref{table:average_Hill} приведено число search trials, которые в среднем требовалось для минимизации функций Shekel и Hill с разной точностью поиска $\epsilon$.
 При этом в скобках в таблице указано число нерешенных задач.
-
 Результаты экспериментов показывают, что при грубой точности решения задачи все методы показывают схожие результаты по числу trials, тогда как при высокой точности решения  алгоритму GSA-DT требуется в 2 раза меньше trials, чем его прототипу.
@@
 \begin{specialtable}[H] 
-	\caption{Среднее число испытаний при минимизации Shekel test functions}\label{table:average_Shekel}
+	\caption{Среднее число испытаний при минимизации Shekel test functions (в скобках указано число нерешенных задач)}\label{table:average_Shekel}
 	\center
@@
 \begin{specialtable}[H] 
-	\caption{Среднее число испытаний при минимизации Hill test functions}\label{table:average_Hill}
+	\caption{Среднее число испытаний при минимизации Hill test functions (в скобках указано число нерешенных задач)}\label{table:average_Hill}
 	\center

### 2021-08-06 2021/ICNAAM/2021_ICNAAM.tex
all the comments are taken into account

@@
 \begin{abstract}
-The paper presents results of solving the problem of simulation and research of processes taking place in a fuel system injector.
+The paper presents results of simulation and optimization of the processes taking place in a fuel system injector.
 Simulation has been carried using tools of the OpenFOAM system.
@@
 can serve as a mathematical model for the process of choosing optimum values of parameters $y=(y_1,y_2,...,y_N)$ of the engineering object under consideration.
-The choice of the value  $y^*$ that determines the best option  is reduced to an approximate (with some required accuracy) solution of the above optimization problem based on the evaluation of object characteristics for the set of options $y^i\in D, \; 1 \leq i \leq k$, determined by available resources.
+Finding the value $y^*$ that determines the best option is reduced to an approximate (with some required accuracy) solution of the above optimization problem based on the evaluation of object characteristics for the set of options $y^i\in D, \; 1 \leq i \leq k$, determined by available resources.
 
 The optimal parameter selection problems under consideration are characterized by the fact that the objective function $\varphi(y)$ is not defined analytically; there is only some algorithm to calculate its values at the points of the domain $D$.
-In this case, one search trial (calculating the function value in a particular point of the search domain) is a computationally intensive operation \cite{Sergeyev1999,Kvasov2013,Kalyulin2017,Sergeyev2018,Paulavicius2020,Lera2021}.
+In this case, one search trial (calculating the function value in a particular point of the search domain) is a computationally intensive operation \cite{Sergeyev1999,Kvasov2013,Kalyulin2017,Sergeyev2018,Paulavicius2020}.%,Lera2021}.
 
@@
 \]
-% Предположение такого рода является достаточно естественным для многих прикладных задач, поскольку относительные вариации целевой функции, характеризующей моделируемую систему, обычно не могут превышать некоторый порог, определяемый ограниченной энергией изменений в системе.
+This kind of assumption is quite natural for many applied problems, since the relative variations of the objective function characterizing the system being modeled usually cannot exceed a certain threshold determined by the limited energy of changes in the system.
 
-At the same time, various dimensionality reduction schemes are widely used to reduce the complexity of designing global optimization algorithms.
+At the same time, various dimensionality reduction schemes are widely used to simplify the rules of global optimization algorithms.
 By means of such schemes, it is possible to reduce solution of multidimensional optimization problems to the solution of one or more nested one-dimensional optimization problems.
@@
 \]
-As follows from the properties of the space-filling curve \cite{Sergeyev2013}, the reduced one-dimensional function $\varphi(y(x))$ satisfies a uniform H{\"o}lder condition
-\[
-\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}, \; x_1,x_2 \in [0,1],
-\]
+As follows from the properties of the space-filling curve \cite{Sergeyev2013}, the reduced one-dimensional function $\varphi(y(x))$ satisfies a uniform H{\"o}lder condition,
+%\[
+%\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}, \; x_1,x_2 \in [0,1],
+%\]
 where the H{\"o}lder constant $H$ is related to the Lipschitz constant $L$ by the relationship $H=2L\sqrt{N+3}$.
@@
 Then new trial is carried out in the interval where the global minimizer of the current minorant is situated and the new trial point $x^{k+1}$ coincides with it.
-
 A detailed description of the algorithm, its various modifications (including parallel ones), and the corresponding convergence  theory are presented in \cite{Strongin2000,Barkalov2010,Strongin2020}.
@@
 \[
-\vec{U}(x,y) = 0,
-\]
-\[
-p(x,y) = 0.
+\vec{U}(x,y) = 0, \; p(x,y) = 0.
 \]
@@
 \[
-\nabla p \cdot \vec{n} = 0,
-\]
-\[
-\vec{U} = 0.
+\nabla p \cdot \vec{n} = 0, \; \vec{U} = 0.
 \]
@@
 \[
-\nabla p \cdot \vec{n} = 0,
-\]
-\[
-\vec{U} = \vec{U_i}.
+\nabla p \cdot \vec{n} = 0, \; \vec{U} = \vec{U_i}.
 \]
@@
 \[
- p = 0,
-\]
-\[
-\vec{U} = \vec{U_o}.
+ p = 0, \vec{U} = \vec{U_o}.
 \]
@@
 The geometry is parametrized by increasing the distance between the red points symmetrically with respect to the channel centreline.
-%Таким образом, в задаче имеется один варьируемый параметр (диаметр сечения канала в его средней части), значения которого изменяются в диапазоне $[0,10]$.
+Thus, the problem has one variable parameter (the diameter of the channel section in its middle part), the values of which vary in the range $[0,10]$.
+The objective function values represent fuel consumption.
 
@@
     \centering
-    \begin{tikzpicture}[scale=0.5,>=stealth]
+    \begin{tikzpicture}[scale=0.45,>=stealth]
          \draw (-5,-2.5,0)--(-15,-2.5,0)--(-15,2.5,0)--(-5,2.5,0.0);%--(0,0.5,0.0)--(5,2.5,0)--(15,2.5,0)--(15,-2.5,0);
@@
 The process of finding the effective channel thickness took 3395 seconds on an Intel(R) Xeon(R) CPU X5670 2.93 GHz, during which time 69 iterations of the global search algorithm were performed.
-% Для решения задачи был использован смешанный локально-глобальный алгоритм \cite{Barkalov2010}, реализованный в Globalizer.
-% При этом в критерии остановки метода было использовано значение $\epsilon = 10^{-3}$.
+The mixed local-global algorithm \cite{Barkalov2010} implemented in Globalizer \cite{globalizerSystem} was used to solve the problem.
+The value $\epsilon = 10^{-3}$ was used in the method’s stopping condition.
 The horizontal component of the velocity at the entrance to the computational domain $\vec{U_i}$  was 10 m/s.
@@
 
-%Найденное решение является удовлетворительным с точки зрения решаемой задачи.
-Например, другие решения, которые получаются при запуске методов локальной оптимизации для решения данной задачи, зависят от выбора стартовой точки.
-И значение целевой функции при этом получается не меньше, чем значение в точке, найденной Globalizer.
+The solution found is satisfactory from the point of view of the problem being solved.
+For example, other solutions that are obtained when running local optimization methods to solve the problem depend on the choice of the starting point.
+The value of the objective function obtained  in these runs is not smaller than the value at the optimum point found by Globalizer.
 

### 2021-08-12 2021/Entropy/entropy.tex
добавил третьего автора

@@
 \newcommand{\orcidauthorB}{0000-0002-8736-0652} % Add \orcidB{} behind the author's name
-%\newcommand{\orcidauthorC}{0000-0002-4013-2329} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorC}{0000-0001-6776-0096} % Add \orcidB{} behind the author's name
 
 % Authors, for the paper (add full first names)
-\Author{Konstantin Barkalov $^{1}$*\orcidA{} and Ilya Lebedev $^{1}$\orcidB{}}
+\Author{Konstantin Barkalov $^{1}$*\orcidA{}, Ilya Lebedev $^{1}$\orcidB{} and Evgeny Kozinov $^{1}$\orcidC{}}
 
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Konstantin Barkalov and Ilya Lebedev}
+\AuthorNames{Konstantin Barkalov, Ilya Lebedev and Evgeny Kozinov}
 
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Barkalov, K.; Lebedev, I.}
+\AuthorCitation{Barkalov, K.; Lebedev, I.; Kozinov, E.}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
@@
 \address[1]{%
-$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhni Novgorod, Russia; ilya.lebedev@itmm.unn.ru (I.L.)}
+$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhni Novgorod, Russia; ilya.lebedev@itmm.unn.ru (I.L.), evgeny.kozinov@itmm.unn.ru (E.A.)}
 %$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Conceptualization and methodology, K.B.; software and validation, I.L.; formal analysis, K.B.; investigation, I.L.; data curation, I.L.; writing---original draft preparation, K.B.; writing---review and editing, K.B.; visualization, I.L.; funding acquisition, K.B.
+\authorcontributions{Conceptualization and methodology, K.B.; software and validation, I.L., E.A.; formal analysis, K.B.; investigation, I.L.; data curation, I.L.; writing---original draft preparation, K.B.; writing---review and editing, K.B.; visualization, I.L.; funding acquisition, K.B.
 All authors have read and agreed to the published version of the manuscript.}

### 2021-08-25 2021/Entropy/entropy.tex
after translation

--- 2021-08-04 2021/Entropy/entropy.tex
@@
 % Full title of the paper (Capitalized)
-\Title{Ускорение метода глобальной оптимизации за счет выделения локальных экстремумов на основе машинного обучения}
-%		   Acceleration of Global Optimization Algorithm by Detecting Local Extrema Based on Machine Learning
+\Title{Acceleration of Global Optimization Algorithm by Detecting Local Extrema Based on Machine Learning}
+		   %Acceleration of Global Optimization Algorithm by Detecting Local Extrema Based on Machine Learning
 
@@
 \address[1]{%
-$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhni Novgorod, Russia; ilya.lebedev@itmm.unn.ru (I.L.)}
+$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhny Novgorod, Russia; ilya.lebedev@itmm.unn.ru (I.L.)}
 %$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
@@
 \abstract{
-В статье рассматриваются задачи глобальной оптимизации и численные методы их решения.
-Задачи указанного класса являются вычислительно-трудоемкими, т.к.
-целевая функция может являться многоэкстремальной, недифференцируемой, и, как правильно, заданной в виде «черного ящика».
-В проведенном исследовании использовался детерминированный алгоритм поиска глобального экстремума, который не основан на идеях мультистарта и nature-inspired algorithms.
-В статье приведены вычислительные правила одномерного алгоритма и nested optimization scheme, которая позволяет применять его для решения сложных многомерных задач.
-Отметим, что сложность решения задач глобальной оптимизации существенным образом зависит от наличия многих локальных экстремумов.
-В работе предложен алгоритм выделения областей притяжения локальных минимумов, основанный на использовании методов машинного обучения (в частности, деревьев решений).
-Использование в выделенных областях алгоритмов локальной оптимизации позволяет значительно ускорить сходимость глобального поиска за счет снижения числа поисковых испытании в окрестностях локальных минимумов.
-Приведены результаты вычислительных экспериментов, проведенных на нескольких сотнях задач глобальной оптимизации разной размерности, которые подтверждают эффект ускорения сходимости (в терминах числа поисковых испытаний, требующихся для решения задачи с заданной точностью).
+The article features the study of global optimization problems and numerical methods of their solution.
+Such tasks are computationally expensive since the objective function can be multi-extremal, nondifferentiable, and, as a rule, given in the form of a "black box".
+The study used a deterministic algorithm for finding the global extremum.
+The said algorithm is based neither on the concept of multistart, nor nature-inspired algorithms.
+The article provides computational rules of the one-dimensional algorithm and the nested optimization scheme which could be applied for solving complex multidimensional problems.
+Note that the solution complexity of global optimization problems essentially depends on the presence of multiple local extrema.
+In this paper, we apply machine learning methods to identify areas of attraction of local minima.
+The use of local optimization algorithms in the selected areas can significantly accelerate the convergence of global search as it could reduce the number of search trials in the vicinity of local minima.
+The results of computational experiments carried out on several hundred global optimization problems of different dimensionalities presented in the paper confirm the effect of accelerated convergence (in terms of the number of search trials required to solve a problem with a given accuracy).
 }
@@
 
-Успешное применение методов машинного обучения (ML) для решения широкого спектра задач приводит к созданию новых подходов к эффективному использованию идей ML в различных областях приложений. 
-Примером класса задач, в которых методы машинного обучения продемонстрировали свою эффективность, являются задачи выявления основных свойств исследуемых явлений (например, физических, экономических или социальных), которые характеризуются стохастической природой или наличием скрытых параметров \cite{Golovenkin2020,Gonoskov2019}.
-Методы машинного обучения успешно используются и для решения сложных задач вычислительной математики, например, для simulation of dynamical systems \cite{Seleznev2019}, решения ordinary, partial or stochastic differential equations \cite{Lagaris1998,Blechschmidt2021,Xu2020} .
+The successful application of machine learning (ML) methods to solve a wide range of problems leads to the emergence of new ways to apply ML for many tasks. 
+Methods of machine learning were shown to be particularly effective for identifying the principal properties of the phenomena (for example, physical, economic, or social), which are stochastic by nature or contain some hidden parameters \cite{Golovenkin2020,Gonoskov2019}.
+ML is also successfully used to solve complex problems of computational mathematics, for example, for simulation of dynamical systems  \cite{Seleznev2019}, solution of ordinary, partial, or stochastic differential equations \cite{Lagaris1998,Blechschmidt2021,Xu2020} .
 
-Одной из таких сложных задач вычислительной математики, к решению которой можно так или иначе применить методы машинного обучения, являются задачи глобальной оптимизации. 
-В таких задачах, как правило, не представляется возможным найти решение аналитически и возникает необходимость построения численных методов для его поиска.
+For example, ML could be applied for solving such a complex problem of computational mathematics as global optimization.
+The solution to this class of problems, as a rule, cannnot be found analytically and, therefore, one needs to construct numerical methods to solve it.
 
-Проблема численного решения задач оптимизации сопряжена со значительными трудностями.
-Во многом они связаны с размерностью и видом целевой функции.
-При этом наиболее сложными являются задачи, в которых целевая функция является многоэкстремальной, недифференцируемой и, более того, заданной в форме черного ящика (т.
-е.
-в виде некоторой вычислительной процедуры, на вход которой подается аргумент, а выходом является соответствующее значение функции).
-Методы решения именно таких задач рассматриваются в данной статье.
+The numerical solution of optimization problems is fraught with significant difficulties.
+In many ways, they are related to the dimensionality and type of the objective function.
+Consequently, the most difficult tasks are those in which the objective function is multi-extremal, nondifferentiable, and, moreover, given in the form of a black box (i.e., in the form of some computational procedure, the input of which is an argument, and the output is the corresponding value of the function).
+These complex problems are the main focus of this article.
 
-Можно выделить несколько подходов к построению численных методов решения задач глобальной оптимизации. 
+There are several approaches to the construction of numerical methods for solving global optimization problems. 
 %Надо ли добавить алгоритмы на основе метамоделей ?
-Ряд алгоритмов основан на идее мультистарта: запуск локального поиска либо начиная с разных стартовых точек, либо с использованием разных параметров.
-Методы локальной оптимизации обладают большой скоростью сходимости.
-При этом одной из основных проблем в мультистартовых схемах является выбор начальных точек, которые бы соответствовали областям притяжения различных локальных решений.
-К решению данной проблемы могут быть успешно применены методы машинного обучения. 
-Например, в \cite{RinnooyKan1987} для выбора перспективных стартовых точек использовались методы кластерного анализа. 
-В \cite{Cassioli2012} область для запуска локального метода выделялась на основе классификации стартовых точке с помощью support vector machine.
+Some algorithms are based on the idea of a multistart: launching a local search either from different starting points or with varying parameters.
+Local optimization methods have a high convergence rate.
+At the same time, one of the main problems in multistart schemes is the choice of starting points that would correspond to the regions of attraction of various local solutions.
+Machine learning methods can be successfully applied to solve this problem.  
+For example, in \cite{RinnooyKan1987} methods of cluster analysis were used to select promising starting points
+In \cite{Cassioli2012} ], the area for launching the local method was allocated based on the classification of starting points using a support vector machine.
 
-Еще одним популярным классом методов, применимым к решению задач глобальной оптимизации, являются метаэвристические алгоритмы. 
-Многие из них основаны на имитации процессов, протекающих в живой природе.
-Для настройки параметров таких алгоритмов также применимы методы машинного обучения.
-Например, в \cite{Jin2005} приведен обзор использования методов машинного обучения в эволюционных алгоритмах.
+Another popular class of methods for solving global optimization problems is metaheuristic algorithms. 
+Many of them are based on imitation of the processes occurring in living nature.
+The parameters of such algorithms could also be tuned using ML.
+For example, \cite{Jin2005} provides an overview of machine learning applications in evolutionary algorithms.
 
-Следует отметить, что алгоритмы последних двух классов не обеспечивают гарантированную сходимость к решению задачи и проигрывают детерминированным алгоритмам по качеству работы \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a certain set).
-Поэтому перспективным является использование именно детерминированных методов.
+Note that the algorithms of the last two classes do not provide guaranteed convergence to the solution of the problem and are inferior to deterministic algorithms in terms of the quality of solution \cite{Kvasov2018,Sergeyev2018} ] (eg, measured by the number of correctly solved problems from a particular set).
+Therefore, deterministic methods seem to be potentially more effective.
 
@@
 
-Для алгоритма глобального поиска в разное время было предложено несколько стратегий ускорения его работы (в терминах числа итераций, требующихся для решения задачи с заданной точностью).
-В данной работе предлагается новый подход к ускорению, основанный на выявления областей притяжения локальных минимумов с использованием методов машинного обучения.
-Выявление областей притяжения и запуск в этих областях локального поиска позволяет существенным образом сократить число испытаний, требующееся методу для достижения глобальной сходимости.
-Сказанное подтверждается результатами экспериментов, проведенных на серии из нескольких сотен тестовых задач.
+Several strategies have been proposed to speed up the operation of the global search algorithm (in terms of the number of iterations required to solve the problem with a given accuracy).
+In this paper, we propose a new approach to acceleration based on identifying areas of attraction of local minima using machine learning methods.
+The identification of regions of attraction and the launch of local search in these regions can significantly reduce the number of trials required for the method to achieve global convergence.
+Experiments carried out on a series of several hundred test problems confirm this statement.
 
@@
 
-В данной работе мы будем рассматривать задачи глобальной оптимизации вида
+In this paper, we will consider global optimization problems of the form
 \begin{eqnarray}\label{main_problem}
@@
 \end{eqnarray}
-Задача (\ref{main_problem}) рассматривается в предположении, что целевая функция является многоэкстремальной, задана как ``черный ящик'', а вычисление ее значений связано с решением задачи численного моделирования и является трудоемкой операцией.
+Problem (\ref{main_problem}) is considered under the assumption that the objective function is multi-extreme, is given in a form of a "black box" , and the calculation of its values is associated with solving the problem of numerical modeling, which makes the solution a labor-intensive operation.
 
-Для многих прикладных задач типичной является ситуация, когда ограниченное изменение вектора параметров $y$ вызывает ограниченное изменение значений $\varphi(y)$.
-Математической моделью, описывающей указанное предположение, является предположение о выполнимости условия Липшица
+A typical situation for many applied problems is when a limited change in the vector of parameters y causes a limited change in the values of $\varphi(y)$.
+The mathematical model describing this premise is based on the assumption that the Lipschitz condition is satisfied
 \[
@@
 \]
-Предположение липшицевости целевой функции типично для многих подходов к разработке оптимизационных алгоритмов \cite{Jones1993,Pinter1996,Zilinskas2008,Evtushenko2009}.
-При этом многие известные методы основаны на различных способах разбиения области поиска на систему подобластей и последующего выбора наиболее перспективной подобласти для размещения очередного испытания (вычисления значения целевой функции) \cite{Jones2009,Zilinskas2010,Evtushenko2013,Kvasov2013,Paulavicius2016}.
+The common assumption for many approaches to the development of optimization algorithms is that the objective function is Lipschitzian  \cite{Jones1993,Pinter1996,Zilinskas2008,Evtushenko2009}.
+At the same time, many known approaches are based on various methods of dividing the search domain into a system of subdomains and then choosing the most promising subdomain for placing the next trial (calculating the value of the objective function)  \cite{Jones2009,Zilinskas2010,Evtushenko2013,Kvasov2013,Paulavicius2016}.
+An important property of global optimization problems is the fact that, in contrast to the problems of finding a local extremum, the global minimum is an integral characteristic of the problem being solved.
+Making sure that the point  $y^*\in D$ is a solution to the problem requires going beyond its neighborhood to the investigation of the entire search domain.
+As a result, when minimizing substantially multi-extremal functions, the numerical method must construct a coverage of the search domain.
+As a result, the number of nodes of this coverage increases exponentially with increasing dimensionality.
+This feature determines the high complexity of solving multiextremal optimization problems making dimensionality a critical factor that affects the complexity of their solution.
 
-Важным свойством задач глобальной оптимизации является тот факт, что в отличие от задач на поиск локального экстремума, глобальный минимум является интегральной характеристикой решаемой задачи.
-Чтобы убедиться, что точка $y^*\in D$ является решением задачи, недостаточно исследовать лишь ее окрестность, требуется исследование всей области поиска.
-Как результат, при минимизации существенно многоэкстремальных функций численный метод должен построить покрытие области поиска и число узлов этого покрытия увеличивается экспоненциально с ростом размерности.
-Эта особенность обуславливает высокую трудоемкость решения задач многоэкстремальной оптимизации, и размерность является решающим фактором, влияющим на сложность их решения.
+The dimensionality in multi-extreme optimization leads to many problems, so, scholars use a wide variety of approaches to reducing it.  
+For example, simplicial or diagonal partition of the search domain allows using methods for solving one-dimensional problems to solve the original multidimensional problem (see, for example,  \cite{PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
+Another well-known approach to dimensionality reduction is using the Peano space-filling curves to map the multidimensional domain onto a one-dimensional interval \cite{Strongin2000}.
 
-Для преодоления сложностей, вызванных размерностью решаемой задачи, в многоэкстремальной оптимизации широко используются различные подходы к уменьшению размерности.  
-Например, simplicial or diagonal partition области поиска позволяет использовать для решения исходной многомерной задачи методы решения одномерных задач (see, for example, \cite{PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
-Известным подходом к редукции размерости является также использование Peano space-filling curves mapping the multidimensional domain onto one-dimensional interval \cite{Strongin2000}.
-
-В данной работе мы будем использовать еще один способ, основанный на the nested optimization scheme \cite{Shi2000,Grishagin2001,VanDam2010,Grishagin2015} and in its generalization \cite{Grishagin2016,Grishagin2016_1}.
-The nested optimization scheme, с одной стороны, не ухудшает свойств минимизируемой функции (в отличие от редукции с использованием Peano curves), а с другой стороны, не требует использования сложных структур данных для поддержки симплексных или диагональных разбиений допустимой области.
-Одновременно с этим схема вложенной оптимизации позволяет свести решение многомерной задачи к решению набора стандартных одномерных задач и использовать широкий спектр алгоритмов одномерной глобальной оптимизации для их решения.
+In this work, we will use another method based on the nested optimization scheme \cite{Shi2000,Grishagin2001,VanDam2010,Grishagin2015} and its generalization \cite{Grishagin2016,Grishagin2016_1}.
+The nested optimization scheme, on the one hand, does not worsen the properties of the minimized function (unlike reduction using Peano curves), and, on the other hand, does not require the use of complex data structures to support simplex or diagonal partitions of the feasible region
+At the same time, the nested optimization scheme makes it possible to reduce the solution of a multidimensional problem to the solution of a set of standard one-dimensional problems, which, then, can be solved by a wide range of one-dimensional global optimization algorithms.
 
@@
 
-В качестве стандартной задачи рассмотрим одномерную задачу многоэкстремальной оптимизации 
+As a standard, let us consider a one-dimensional multiextremal optimization problem
 \begin{equation}\label{uni_problem}
@@
 \end{equation}
-с целевой функцией, удовлетворяющей условию Липшица.
+with an objective function satisfying the Lipschitz condition
 
-Приведем описание global search algorithm (GSA) для решения базовой задачи в соответствии с \cite{Strongin2000}.
-В процессе своей работы GSA порождает последовательность точек $x^i$, в которых вычисляются значения целевой функции $z^i=\varphi(x^i)$. 
-Будем называть процесс вычисления значения целевой функции \textit{trial}.
+Here is the description of global search algorithm (GSA) for solving the basic problem in accordance with \cite{Strongin2000}.
+]. In the course of its work, GSA generates a sequence of points  $x^i$, at which the values of the objective function  $z^i=\varphi(x^i)$ are calculated. 
+We will refer to the process of calculating the value of the objective function as  \textit{trial}.
 
-В соответствии с алгоритмом первые два испытания проводятся в граничных точках отрезка $[0,1]$, т.е. $x^0=0,\;x^1=1$. 
-В этих точках вычисляются значения целевой функции $z^0=\varphi(x^0),\;z^1=\varphi(x^1)$ и устанавливается значение счетчика $k=1$. 
-Точка очередного испытания $x^{k+1}, k\geq 1,$ выбирается в соответствии со следующими действиями.
+In accordance with the algorithm, the first two trials are carried out at the boundary points of the segment $[0,1]$, т.е. $x^0=0,\;x^1=1$. 
+At these points, the values of the objective function  $z^0=\varphi(x^0),\;z^1=\varphi(x^1)$ и уare calculated and the counter value is set to $k=1$. 
+The point of the next trial $x^{k+1}, k\geq 1,$ is selected in accordance with the following procedure:
 
- Step 1. Перенумеровать нижним индексом (начиная с 0) точки $x^i,\:0\leq i\leq k$, проведенных испытаний в порядке возрастания координаты, т.е.
+ Step 1. Reindex  (starting at 0) the points xl, $x^i,\:0\leq i\leq k$, of the trials conducted  in ascending order of the coordinate, i.e.
 \begin{equation}\label{xt}
@@
 \end{equation} 
-Сопоставить точкам $x_i, \; 0\leq i\leq k$, вычисленные в них значения целевой функции $z_i=\varphi(x_i), \; 0\leq i\leq k$.
+Compare the values of the objective function $z_i=\varphi(x_i), \; 0\leq i\leq k$ to the points $x_i, \; 0\leq i\leq k$, at which these values were calculated.
 
-Step 2. Вычислить максимальное абсолютное значение относительной первой разности
+Step 2. Calculate the maximum absolute value of the relative first difference
 \begin{equation}\label{mu}
@@
 \end{equation}
-где $\Delta_i = x_i-x_{i-1}$. Если вычисленное в соответствии с данной формулой значение равно нулю, то положить $\mu = 1$.
+where $\Delta_i = x_i-x_{i-1}$. If the value calculated in accordance with this formula is equal to zero, then take $\mu = 1$.
 
-Step 3. Для всех интервалов $(x_{i-1},x_i),1\leq i\leq k$,  вычислить значение
+Step 3. For all the intervals $(x_{i-1},x_i),1\leq i\leq k$,  calculate the value
 \begin{equation}\label{R}
@@
 \end{equation} 
-называемое \textit{характеристикой} интервала; величина $r>1$ является параметром алгоритма.
+refered to as the \textit{characteristics} of the interval; value $r>1$ is the parameter of the algorithm.
 
-Step 4. Найти интервал $(x_{t-1},x_t)$ с максимальной характеристикой
+Step 4. Find interval $(x_{t-1},x_t)$ with the maximum characteristics
 \begin{equation}\label{MaxR}
@@
 \end{equation}
-Если максимальная характеристика соответствует нескольким интервалам, то в качестве $t$ выбрать минимальное число, удовлетворяющее (\ref{MaxR}).
+If the maximum characteristics corresponds to several intervals, then choose the minimum number that satisfies as $t$ (\ref{MaxR}).
 
-Step 5. Провести новое испытание в точке
+Step 5. Carry out a new trial at the point
 \begin{equation}\label{xk1}
@@
 
-Алгоритм прекращает свою работу при выполнении условия $\Delta_t<\epsilon$; здесь $t$ из (\ref{MaxR}), а $\epsilon>0$ есть заданная точность. 
-В качестве оценки решения задачи выбираются значения 
+The algorithm stops when the condition  $\Delta_t<\epsilon$ is satisfied; here $t$ from (\ref{MaxR}), and $\epsilon>0$ is a given accuracy. 
+The following values are selected to evaluate the efficiency of the solution
 \[
@@
 
-Теоретические условия, определяющие сходимость алгоритма, представлены в \cite{Strongin2000}. 
-Работа алгоритма при минимизации конкретной многоэкстремальной функции, которая задаются в соответствии с формулой (\ref{hill}), проиллюстрирована на figure \ref{fig1}.
-При запуске алгоритма использовался параметр $r=2.2$ из (\ref{R}) и значение $\epsilon = 10^{-3}$ в критерии остановки метода.
-На рисунке \ref{fig1} изображен график objective function и точки 71 поискового испытания, потребовавшихся GSA для решения задачи с указанной точностью.
-Наглядно видна проблема всех методов глобальной оптимизации -- концентрация точек в окрестностях локальных минимумов задачи, которые не являются глобальным решением.
+The theoretical conditions that determine the convergence of the algorithm are presented in \cite{Strongin2000}. 
+The operation of the algorithm during the minimization of a specific multiextremal function, which is specified in accordance with formula  (\ref{hill}), is shown in Figure \ref{fig1}.
+The algorithm was launched with the parameter $r=2.2$ from (\ref{R}) and value $\epsilon = 10^{-3}$in the criterion for stopping the method.
+Figure \ref{fig1} shows the objective function graph and points of 71 search trials which GSA needed to solve the problem to the specified accuracy.
+This highlights the problem of all methods of global optimization - the concentration of points in the vicinity of local minima of the problem, which are not a global solution.
 
@@
 \includegraphics[width=1.0\linewidth]{HillAGP90.png}
-\caption{Минимизация функции вида (\ref{hill}) с помощью GSA.}
+\caption{Minimization of a function of the form  (\ref{hill}) using GSA.}
 \label{fig1}
@@
 
-\subsection{Machine Learning Regression как инструмент для выделения областей притяжения локальных экстремумов}\label{TreeGSA}
+\subsection{Machine Learning Regression as a tool for identifying areas of attraction of local extrema}\label{TreeGSA}
 
-Функции, рассматриваемые в рамках данного исследования, принадлежат классу липшицевых функций.
-Поэтому классические методы регрессии (например, полиномиальная регрессия, когда функция приближается полиномом заданной степени) не будут хорошо соответствовать поведению функции.
+The functions considered in this study belong to the class of Lipschitzian functions.
+Therefore, classical regression methods (for example, polynomial regression, where a function is approximated by a polynomial of a given degree) will not properly match the behavior of the function.
+A more powerful tool for this task is regression splines.
+When constructing a regression spline, the domain of definition is divided into K non-overlapping subdomains.
+In each of such domains, the function is approximated by a polynomial.
+Dividing the interval into a sufficient number of subdomains allows one to very accurately approximate the original function.
 
-Более мощным инструментом являются регрессионные сплайны.
-При построении регрессионного сплайна область определения разбивается на $K$ непересекающихся подобластей, в каждой из которых функция аппроксимируется полиномом.
-При разбиении интервала на достаточное число подобластей позволяет очень точно аппроксимировать исходную функцию.
+Regression can also be constructed using such a powerful tool as artificial neural networks. 
+Different types of networks can be used to build the regression, for example, multilayer perceptron, radial basis function network, etc.
+However, in both of these cases, the model itself (spline or neural network) becomes rather complex for the analysis required to solve the given problem (identifying areas of attraction of local extrema).
 
-Построить регрессию можно также и с использованием такого мощного инструмента, как искусственные нейронные сети. 
-Для построения регрессии могут быть использованы сети разного типа, например, multilayer perceptron, radial basis function network, и т.д. 
-Однако в обоих этих случаях модель (сплайн или же нейронная сеть) сама становится сложной для того анализа, который требуется выполнить в решаемой проблеме (выделение областей притяжения локальных экстремумов).
+Therefore, within the framework of the study, we chose a regression model based on decision trees to analyze the local behavior of a function.
+For example, if the objective function is properly approximated by polynomials, then the polynomial regression, of course, will appropriately convey the properties of the function.
+However, if there is a more complex relationship between them, then the decision tree can surpass the classical variants of regression in terms of the quality of the approximation.
+%Здесь нужен рисунок - функция Шекеля, и два варианта регрессии - полином и деревья решений. Шекель должен плохо полиномами представляться, т.к. там функции вида 1/x
+At the same time, the regression based on the decision tree makes it possible to easily identify the areas of attraction of local extrema with sufficient accuracy.
 
-Поэтому в рамках проведенного исследования для анализа локального поведения функции мы выбрали регрессионную модель, основанную на деревьях решений. 
-Например, если целевая функция хорошо аппроксимируется полиномами, то полиномиальная регрессия, конечно же, будет адекватно передавать свойства функции.
-Однако если между имеет место более сложная зависимость, тогда дерево решений может превзойти по качеству аппроксимации классические варианты регрессии.
-%Здесь нужен рисунок - функция Шекеля, и два варианта регрессии - полином и деревья решений. Шекель должен плохо полиномами представляться, т.к. там функции вида 1/x
-Одновременно с этим регрессия на основе дерева решений позволяет с достаточной точностью легко выделять области притяжения локальных экстремумов.
-
-Построение регрессии с помощью дерева решений включает два основных этапа:
+Building a regression using a decision tree consists of two main steps:
 \begin{itemize}
-	\item Область поиска $D$ разбивается на $J$ непересекающихся подобластей $D_1, D_2, ..., D_J$, причем $D = \bigcup_{j=1}^{J}{D_j}$.
-	\item Любому значению, попадающему в область $D_j$, т.е. $x \in D_j$, ставится в соответствие среднее значение $c_j$ по значениям training trials, попавших в эту область.
+	\item Search domain $D$ is divided into $J$ non-overlapping subdomains  $D_1, D_2, ..., D_J$, provided that $D = \bigcup_{j=1}^{J}{D_j}$.
+	\item Any value falling into the subdomain $D_j$, i.e. $x \in D_j$, is matched to the average value  $c_j$ based on the training trial values that fall into this subdomain. 
 \end{itemize}
 
-Фактически, деревья решений строят модель функции вида
+In fact, decision trees build a model of a function of the form
 \begin{equation}\label{dtree}
@@
 
-Указанная модель является, с одной стороны, довольно простой, а с другой стороны, адекватно отражает интересующие нас свойства функции (наличие или отсутствие локальных минимумов).
-Выделение областей притяжения локальных минимумов с использованием модели (\ref{dtree}) можно организовать следующим образом.
+This model is, on the one hand, rather simple, and on the other hand, it adequately reflects the properties of the function under investigation (the presence or absence of local minima).
+The selection of regions of attraction of local minima using model (\ref{dtree}) can be organized as follows.
 
-Пусть $x^{k+1}$ -- точка текущего испытания. Для данной точки ищется номер $j$, такой что $x^{k+1} \in D_j$. 
-Далее сравниваются значения $c_j$, соответствующие соседним подобластям. Если выполняется одно из условий
+Let $x^{k+1}$ be the point of the current trial. For a given point, an index  $j$ is sought such that $x^{k+1} \in D_j$. 
+Next, the  $c_j$ values corresponding to neighboring subdomains are compared. If one of the conditions is met
 \begin{gather}\label{LocalPointC}
@@
 \end{gather}
-то область $D_j$ считается областью притяжения локального минимума. В ней можно запускать локальный поиск, а впоследствии -- исключить данную подобласть из глобального поиска.
-В качестве локального алгоритма может использоваться любой из локальных методов нулевого порядка, например,  golden section search or parabolic interpolation \cite{Press}.
+then the subdomain $D_j$ is considered the area of attraction of the local minimum. Here, one can start a local search, and subsequently exclude this subdomain from the global search. 
+Any of the zero-order local methods can be used as a local algorithm, for example, golden section search or parabolic interpolation \cite{Press}.
 
-Для того, чтобы модифицировать алгоритм глобального поиска из subsection \ref{CoreGSA} с целью исключения областей притяжения локальных минимумов, сопоставим каждой точке испытания $x^i$, получаемой при работе алгоритма, дополнительный признак $q^i \in \{0,1,2\}$, который будет характеризовать свойства данной точки.
-Значение $q^i=0$ присваивается по умолчанию и означает что точка $x^i$ получена в результате работы правила (\ref{xk1}) алгоритма глобального поиска.
-Значение $q^i=1$ присваивается точкам, полученным в результате работы локального метода, при этом значение $q^i=2$ соответствует точке локального минимума, найденной в результате работы локального поиска.
+In order to modify the global search algorithm from subsection  \ref{CoreGSA} to exclude the regions of attraction of local minima, we associate each test point  $x^i$, obtained during the operation of the algorithm with an additional criterion $q^i \in \{0,1,2\}$, which will characterize the properties of this point.
+The value  $q^i=0$ is assigned by default and indicates that point $x^i$ is obtained as a result of rule (\ref{xk1}) of the global search algorithm.
+The value $q^i=1$ is assigned to the points obtained as a result of the work of the local method, while the value  $q^i=2$ corresponds to the point of the local minimum found as a result of the work of the local search.
 
-Напомним, что верхний индекс соответствует номеру итерации, на которой было проведено испытание в данной точке, а нижний индекс -- номеру точки в ряду (\ref{xt}).
+Recall that the superscript corresponds to the number of the iteration at which the trial was carried out at this point, and the subscript corresponds to the number of the point in the row  (\ref{xt}).
 
-Опишем теперь модифицированный алгоритм глобального поиска, использующий деревья решений для выделения и исключения областей притяжения локальных минимумов; будем в дальнейшем ссылаться на данный алгоритм как на GSA-DT.
+Let us now describe a modified global search algorithm that uses decision trees to isolate and exclude areas of attraction of local minima; we will further refer to this algorithm as GSA-DT.
 
-Шаги 1--5 алгоритма GSA-DT полностью повторяют шаги 1--5 .
+Steps 1-5 of the GSA-DT algorithm are the same as steps 1-5.
 
-Step 6. Если $q_{t-1} \in \{1,2\}$ или $q_t \in \{1,2\}$, то перейти к проверке критерия остановки. Иначе перейти на Step 7.
+Step 6. If $q_{t-1} \in \{1,2\}$ or $q_t \in \{1,2\}$, then go to check the stopping criterion. Otherwise, go to Step 7.
 
 Step 7.
-Построить дерево решений на основе результатов выполненных испытаний, и получить соответствующую кусочно линейную аппроксимацию, которая каждой подобласти $D_1, D_2, ..., D_J$ сопоставляет значение $c_1, c_2, ..., c_J$.
+Construct a decision tree based on the results of the tests performed, and obtain the corresponding piecewise linear approximation, which assigns the value $c_1, c_2, ..., c_J$ to each subdomain  $D_1, D_2, ..., D_J$.
 
 Step 8.
-Для точки $x^{k+1}$ текущего испытания найти номер $j$ такой, что $x^{k+1} \in D_j$ и проверить выполнимость условия (\ref{LocalPointC}).
-Если условие (\ref{LocalPointC}) выполняется, запустить в области $D_j$ локальный поиск из точки $x^{k+1}$.
-Результаты всех испытаний, выполненных в процессе работы локального поиска в точках $x^{k+2}, ...,x^{k+k_{local}}$, сохраняются в информационной базе алгоритма и используются на последующих итерациях.
-Все эти точки получают признак $q^i=1$, $i = k+2, ... , k+k_{local}$. Точке, соответствующей найденному локальному минимуму, присваивается признак равный 2.
+For the point $x^{k+1}$ of the current trial, find a number $j$ such that $x^{k+1} \in D_j$ and check whether the condition  (\ref{LocalPointC}) is satisfied.If condition (\ref{LocalPointC}) is satisfied, start a local search in the domain $D_j$ from the point  $x^{k+1}$.
+The results of all trials performed during the local search at the points $x^{k+2}, ...,x^{k+k_{local}}$ are stored in the information base of the algorithm and are used at subsequent iterations. 
+All these points receive the attribute  $q^i=1$, $i = k+2, ... , k+k_{local}$. The attribute equal to 2 is assigned to the point corresponding to the found local minimum.
 
-Критерий остановки модифицированного алгоритма будет выглядеть следующим образом. 
-Алгоритм прекращает свою работу при выполнении одного из условий:
+The stopping criterion of the modified algorithm will look as follows. 
+The algorithm stops when one of the following conditions is met:
 
 \begin{itemize}
-	\item $|x_{t} - x_{t-1}|<\epsilon$, где $t$ из (\ref{MaxR}), а $\epsilon>0$ есть заданная точность;
-	\item $|x^{k+1} - x_{t-1}|<\epsilon$ и $q_{t-1} = 2$,
-	\item $|x^{k+1}-x_{t}|<\epsilon$ и $q_{t} = 2$.
+	\item $|x_{t} - x_{t-1}|<\epsilon$, where $t$ from (\ref{MaxR}), and $\epsilon>0$ is a given accuracy;
+	\item $|x^{k+1} - x_{t-1}|<\epsilon$ and $q_{t-1} = 2$,
+	\item $|x^{k+1}-x_{t}|<\epsilon$ and $q_{t} = 2$.
 \end{itemize}
 
-Для примера рассмотрим работу алгоритма GSA-DT при минимизации той же многоэкстремальной функции, что представлена на fugire \ref{fig1}.
-При запуске алгоритма использовались такре же параметры: параметр $r=2.2$ из (\ref{R}) и значение $\epsilon = 10^{-3}$ в критерии остановки метода.
-Рисунок \ref{fig2} иллюстрирует работу алгоритма GSA-DT.
-Кроме графика целевой функции на нем изображена кусочно-постоянная аппроксимация вида (\ref{dtree}), построенная на финальной стадии поиска.
-Черные точки на графике соответствуют фазе глобального поиска, зеленые точки соответствуют работе локального метода.
-Всего методу GSA-DT потребовалось 49 испытаний для решения задачи; накопление точек испытаний в окрестностях локальных минимумов отсутствует.
+For example, consider the operation of the GSA-DT algorithm with minimization of the same multi-extreme function, which is presented  Figure \ref{fig1}.
+The same parameters were used at the start of the algorithm: the parameter  $r=2.2$ from (\ref{R}) and value $\epsilon = 10^{-3}$ in the stopping criterion of the method.
+Figure \ref{fig2} illustrates the operation of the GSA-DT algorithm.
+In addition to the graph of the objective function, it shows a piecewise constant approximation of the form (\ref{dtree}) built at the final stage of the search.
+Black points on the graph correspond to the global search phase, green points correspond to the work of the local method.
+In total, the GSA-DT method required 49 trials to solve the problem; there was no accumulation of trial points in the vicinity of local minima.
 
@@
 \includegraphics[width=1.0\linewidth]{HillTree90.png}
-\caption{Минимизация функции вида (\ref{hill}) с помощью GSA-DT.}
+\caption{Using GSA-DT to minimize the function (\ref{hill}) }
 \label{fig2}
@@
 
-\subsection{Адаптивная схема редукции размерности }
+\subsection{Adaptive Dimension Reduction Scheme}
 
-Рекурсивная схема вложенной оптимизации основана на известном соотношении \cite{Grishagin2001} 
+The recursive nested optimization scheme is based on the well-known relation  \cite{Grishagin2001} 
 \begin{equation}\label{nested}
@@
 \end{equation}
-которое позволяет свести решение исходной многомерной задачи (\ref{main_problem}) к решению семейства рекурсивно связанных одномерных подзадач.
+which allows reducing the solution of the original multidimensional problem (\ref{main_problem}) to the solution of a family of recursively connected one-dimensional subproblems.
 
-Для формального описания nested optimization scheme введем семейство функций, определяемых в соответствии с соотношениями 
+For a formal description of the nested optimization scheme, we introduce a family of functions defined in accordance with the relations
 \begin{equation}\label{nested_N}
@@
 
-Тогда, в соответствии с (\ref{nested}), для решения многомерной задачи (\ref{main_problem}) достаточно решить одномерную задачу  
+Then, in accordance with  (\ref{nested}), solving the multidimensional problem  (\ref{main_problem}) is reduced to solving a one-dimensional problem 
 \begin{equation}\label{nested_1}
@@
 \end{equation}
-Однако каждое вычисление значения функции $\varphi_1$ в некоторой фиксированной точке $y_1$ предполагает решение одномерной задачи оптимизации второго уровня 
+However, each calculation of the value of the function $y_1$ at some fixed point y1 presupposes the solution of the one-dimensional optimization problem of the second level
 \begin{equation}
@@
 \end{equation}
-Вычисление значений функции $\varphi_2$ в свою очередь требует одномерной минимизации функции $\varphi_3$ и т.д. вплоть до решения задачи
+Calculation of the values of the function  $\varphi_2$ in turn, requires one-dimensional minimization of the function  $\varphi_3$ all the way to the solution of the problem
 \begin{equation}
@@
 \end{equation}
-на последнем уровне рекурсии.
+at the last level of recursion.
 
-Решение возникающего в схеме вложенной оптимизации множества подзадач (\ref{nested_i}) может быть организовано различными способами. 
-Очевидный способ (детально проработанный в \cite{Grishagin2001,Grishagin2015} основан на решении подзадач в соответствии с рекурсивным порядком их порождения.
-Однако здесь возникает потеря значительной части информации о целевой функции.
+The solution of the set of subtasks arising in the nested optimization scheme  (\ref{nested_i}) can be organized in different ways. 
+The obvious way (described in detail in \cite{Grishagin2001,Grishagin2015}) is based on solving subproblems in accordance with the recursive order of their generation.
+However, a significant part of the information about the objective function is lost here.
 
-Иным  подходом является адаптивная схема, в которой все подзадачи решаются одновременно, что позволяет более полно учитывать информацию о многомерной задаче и за счет этого ускорять процесс ее решения.
-Данный подход был теоретически обоснован и апробирован в \cite{Grishagin2016,Grishagin2016_1,Grishagin2018}.
+Another approach is an adaptive scheme, in which all subtasks are solved simultaneously, which allows taking into account much more information about a multidimensional problem, thereby speeding up the process of its solution.
+This approach was theoretically substantiated and tested in  \cite{Grishagin2016,Grishagin2016_1,Grishagin2018}.
 
-Кратко отметим, что в рамках исходной схемы вложенной оптимизации порождаемые подзадачи решаются строго последовательно; получаемая в результате иерархическая схема порождения и решения подзадач имеет вид дерева.
-Построение этого дерева происходит динамически в процессе решения исходной задачи (\ref{main_problem}).
-При этом вычисление одного значения функции $\varphi_i(y_1,y_2,...,y_i)$ на $i$-м уровне требует полного решения всех задач одного из поддеревьев уровня $i+1$.
+Note that within the framework of the original nested optimization scheme, the generated subproblems are solved only sequentially; the resulting hierarchical scheme for generating and solving subproblems has the form of a tree.
+The construction of this tree occurs dynamically in the process of solving the original problem (\ref{main_problem}).
+In this case, the calculation of one value of the function $\varphi_i(y_1,y_2,...,y_i)$ at the $i$-th ith level requires a complete solution of all problems of one of the subtrees of level  $i+1$.
 
-Adaptive nested optimization scheme  редукции размерности изменяет порядок решения подзадач: они будут решаться не по одной (в соответствии с их иерархией в дереве задач), а одновременно, т.е.
-будет существовать некоторое множество подзадач, находящихся в процессе решения.
-В рамках адаптивной схемы:
+The adaptive nested optimization scheme of dimensionality reduction changes the order of solving subproblems: they will be solved not one by one (in accordance with their hierarchy in the problem tree), but simultaneously, i.e.
+there will be a set of subtasks in the process of solution.
+Within the adaptive scheme:
 \begin{itemize}
 	\item 
-для вычисления значения функции $i$-го уровня из (\ref{nested_i}) порождается новая задача уровня $i+1$, в которой проводится только одно испытание, после чего новая порожденная задача включается в множество уже имеющихся задач, подлежащих решению;
+to calculate the value of $i$-th level function from (\ref{nested_i}) a new $i+1$ level problem is generated, in which only one trial is carried out, after which the new generated problem is included in the set of already existing problems to be solved;
 	\item 
-	итерация глобального поиска состоит в выборе одной (наиболее перспективной) задачи из множества имеющихся задач, в которой проводится одно испытание; точка проведения нового испытания определяется в соответствии с базовым алгоритмом глобального поиска из subsection \ref{CoreGSA} или же модифицированным алгоритмом из subsection \ref{TreeGSA};
+	iteration of the global search consists in choosing one (most promising) problem from the set of available problems, in which one trial is carried out; the new trial point is determined according to the basic global search algorithm from subsection \ref{CoreGSA} or a modified algorithm from subsection  \ref{TreeGSA};
 	\item
-в качестве минимальных значений функций из (\ref{nested_i}) используются их текущие оценки, полученные на основе накопленной поисковой информации.
+the minimum values of functions from  (\ref{nested_i}) are their current estimates obtained based on accumulated search information.
 \end{itemize}
 
-Краткое описание основных шагов адаптивной схемы редукции размерности состоит в следующем.
+A brief description of the main steps of the adaptive dimensionality reduction scheme is as follows.
 
-Пусть вложенные подзадачи вида (\ref{nested_i}) решаются с помощью алгоритма глобального поиска, описанного в subsection \ref{CoreGSA}.
-Тогда каждой подзадаче (\ref{nested_i}) можно присвоить некоторое числовое значение, называемое характеристикой этой задачи.
-В качестве такой характеристики можно взять значение $R(t)$ из (\ref{MaxR}), т.е.
-максимальную характеристику из характеристик интервалов, сформированных в данной задаче.
-Чем выше значение данной характеристики, тем более перспективной является подзадача для продолжения поиска в ней глобального минимума исходной задачи (\ref{main_problem}).
-Поэтому на каждой итерации выбирается подзадача с максимальной характеристикой для проведения в ней очередного испытания.
-Это испытание либо приводит к вычислению значения целевой функции $\varphi(y)$ (если выбранная подзадача принадлежала уровню $j=M$), либо порождает новые подзадачи согласно (\ref{nested_i}) при $j\leq M-1$.
-В последнем случае новые порожденные задачи добавляются к текущему множеству задач, вычисляются их характеристики и процесс повторяется.
-Завершение процесса оптимизации происходит, когда в корневой задаче выполняется условие остановки алгоритма, решающего эту задачу.
+Let nested subproblems of the form (\ref{nested_i}) be solved using the global search algorithm described in subsection  \ref{CoreGSA}.
+Then each subproblem  (\ref{nested_i})  can be assigned some numerical value called a characteristic of this problem.
+As such a characteristic, we can take the value $R(t)$ from (\ref{MaxR}), i.e.
+the maximum characteristic from the characteristics of the intervals formed in this task.
+The higher the values of this characteristic, the more promising is the search for the global minimum of the original problem in the given subproblem (\ref{main_problem}).Therefore, at each iteration, a subproblem with the maximum characteristic is selected, and a trial is carried out in it.
+This either leads to the calculation of the value of the objective function  $\varphi(y)$(if the selected subtask belonged to the level  $j=M$), or generates new subproblems according to  (\ref{nested_i}) for $j\leq M-1$.
+In the latter case, new generated tasks are added to the current set of tasks, their characteristics are calculated, and the process is repeated.
+The completion of the optimization process occurs when the condition for stopping the algorithm that solves this problem is satisfied for the root problem.
 
@@
 
-Numerical experiments were performed on the Lobachevsky supercomputer of the University of Nizhni Novgorod (operating system -- CentOS 6.4, management system -- SLURM).
+Numerical experiments were performed on the Lobachevsky supercomputer of the University of Nizhny Novgorod (operating system CentOS 7.2, management system SLURM).
 One supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM.
@@
 a total of 16 CPU cores are available on the node).
+All the algorithms were implemented in Cd— +; GCC 5.5.0 was used for compilation on the supercomputer.
 
-Традиционный подход к оценке эффективности методов глобальной оптимизации основан на численном решении этими методами серии задач из некоторого класса. 
-При этом подразумевается, что очередная решаемая задача генерируется посредством некоторого алгоритма.
-Типичными примерами таких классов тестовых функций являются функции  Шекеля и Хилла. 
-Первый из них (обозначит его $F_{SH}$) основан на формуле 
+The traditional approach to assessing the effectiveness of global optimization methods is based on using these methods to find the numerical solution of a series of a certain class of problems. 
+In this case, the assumption is that a certain algorithm is used to generate the next problem to be solved. 
+Typical examples of such test function classes are Shekel and Hill functions. 
+The first of them (denoted $F_{SH}$) is based on the formula.
 \begin{equation}\label{shekel}
@@
 \end{equation}
-где параметры $1\le K_j\le 3,\: 0 < A_j,\: C_j < 10, \;$ являются независимыми случайными величинами, равномерно распределенными в указанных интервалах.
-Следующий генератор (обозначим его $F_{HL}$) определяется выражением
+where parameters $1\le K_j\le 3,\: 0 < A_j,\: C_j < 10, \;$ are independent random variables uniformly distributed in the indicated intervals. 
+The next generator (denoted   $F_{HL}$) is determined by the expression
 \begin{equation}\label{hill}
@@
 \end{equation}
-где значения параметров  $A_j,\: B_j,\: 1 \le j \le 14$, независимо и равномерно распределены в интервале $[-1 ,1 ]$.
+where the values of the parameters $A_j,\: B_j,\: 1 \le j \le 14$, are independently and uniformly distributed in the interval  $[-1,1]$.
 
-Проведем сравнение базового global search algorithm (GSA) и его модификации, использующей деревья решений (GSA-DT), c известным алгоритмом глобальной оптимизации DIRECT \cite{Jones2009}.
-Сравнение методов будем проводить при решении 1000 задач из классов $F_{SH}$ и $F_{HL}$.
+Let us compare the basic global search algorithm (GSA) and its decision tree-based modification (GSA-DT) with the well-known DIRECT global optimization algorithm  \cite{Jones2009}.  
+The decision trees in the GSA-DT algorithm were built using the OpenCV 4.5.1 library (class cv :: ml :: DTrees).
+A regression function was constructed by a single decision tree.
+This allowed us to obtain a piecewise constant approximation of the objective function in which one or several trial points corresponded to each leaf node of the tree.
+The tree was built without any limitations on maximum depth (MaxDepth); the accuracy of tree construction (RegressionAccuracy) was 10-3; the minimum number of trial points in the tree node (MinSampleCount) was equal to one; all other parameters were set by default.
+Stopping tree building occurred if all absolute differences between an estimated value in a node and values of train samples in this node are less than accuracy.
 
-Задачу будем считать решенной корректно в случае, если после остановки метода по точности (т.е.
-когда длина текущего поискового интервала станет меньше $\epsilon$) текущая оценка оптимума $x_k^*$ будет лежать в $\epsilon$-окрестности известного решения задачи $x^*$, т.е.
-будет выполняться условие $|x^*-x_k^*| \leq \epsilon$.
+The global optimization methods discussed above were compared when solving 100 problems from the  $F_{SH}$ and $F_{HL}$ classes.
+The problem will be considered correctly solved if after stopping the method by accuracy (i.e.
+when the length of the current search interval becomes less than $\epsilon$) the current estimate of the optimum  $x_k^*$ lies in the $\epsilon$-neighborhood of the known solution of the problem  $x^*$, *, i.e.
+if the condition  $|x^*-x_k^*| \leq \epsilon$ is satisfied.
 
-В таблицах \ref{table:average_Shekel} и \ref{table:average_Hill} приведено число поисковых испытаний, которые в среднем требовалось для минимизации функций Шекеля и Хилла с разной точностью поиска.
-При этом в скобках в таблице указано число нерешенных задач.
-
-Результаты экспериментов показывают, что при грубой точности решения задачи все методы показывают схожие результаты по числу trials, тогда как при высокой точности решения  алгоритму GSA-DT требуется в 2 раза меньше trials, чем его прототипу.
-Одновременно GSA-DT превосходит метод DIRECT как по среднему числу search trials, так и по числу корректно решенных задач.
-В частности, если использовать точность $\epsilon = 10^{-2}$, то метод DIRECT останавливается слишком рано, и не находит глобального решения во многих задачах.
-Поэтому в дальнейших экспериментах, в которых решаются многомерные задачи, мы не будем использовать DIRECT для сравнения, т.к.
-при решении многомерных задач с остановкой по точности данный метод обеспечиват корректное решение не более чем $50\%$ задач.
+Tables \ref{table:average_Shekel} and \ref{table:average_Hill} show the number of search trials that, on average, were required to minimize the Shekel and Hill functions with different search accuracy  $\epsilon$.
+The number of unsolved problems is indicated in parentheses.
+The experimental results show that with a rough solution to the problem, all methods show similar results in terms of the number of trials, while with a high solution accuracy, the GSA-DT algorithm requires 2 times fewer trials than its prototype.
+At the same time, GSA-DT outperforms the DIRECT method both in the average number of search trials and in the number of correctly solved problems.
+In particular, if we use the accuracy $\epsilon = 10^{-2}$, then the DIRECT method stops too early and does not find a global solution to many problems.
+Therefore, in further experiments in which multidimensional problems are solved, we will not use DIRECT for comparison, since when solving multidimensional problems with stopping by accuracy, this method will provide a correct solution to no more than  $50\%$ of problems.
 
 \begin{specialtable}[H] 
-	\caption{Среднее число испытаний при минимизации Shekel test functions}\label{table:average_Shekel}
+	\caption{The average number of tests when minimizing Shekel test functions (the number of unsolved problems is indicated in parentheses))}\label{table:average_Shekel}
 	\center
@@
 \begin{specialtable}[H] 
-	\caption{Среднее число испытаний при минимизации Hill test functions}\label{table:average_Hill}
+	\caption{The average number of tests when minimizing Hill test functions (the number of unsolved problems is indicated in parentheses))}\label{table:average_Hill}
 	\center
@@
 
-В следующей серии экспериментов решались многомерные задачи. 
-Известным генератором тестовых задач многоэкстремальной оптимизации является генератор GLKS \cite{Gaviano2003}.
-С его помощью можно порождать тестовые функции с заданными свойствами: числом локальных экстремумов, их областями притяжения, точкой глобального минимума и значением целевой функции в ней, и т.д.
-Процедура генерирования тестовых функций состоит в переопределении выпуклой квадратичной функции (параболоида) при помощи полиномов. Тестовые функции определяются пятью параметрами:
+The next series of experiments involved the solution of multidimensional problems. 
+A well-known generator of multi-extremal optimization test problems is GLKS  \cite{Gaviano2003}.
+It can be used to generate test functions with given properties: the number of local extrema, their areas of attraction, the global minimum point, and the value of the objective function at this point, etc.
+The procedure for generating test functions is based on using polynomials to redefine a convex quadratic function (paraboloid).
+Trial functions are defined by five parameters:
 \begin{itemize}
-	\item размерность задачи $N$;
-	\item число локальных минимумов $l$;
-	\item значение глобального минимума $f^*$;
-	\item радиус области притяжения global optimizer $\rho^*$;
-	\item расстояние между global optimizer и вершиной параболоида $d^*$.
+	\item dimensionality of the problem  $N$;
+	\item the number of local minima is $l$;
+	\item value of the global minimum $f^*$;
+	\item radius of the area of attraction of the global optimizer $\rho^*$;
+	\item рthe distance between the global optimizer and the vertex of the paraboloid $d^*$.
 \end{itemize}
-Изменяя указанные параметры, можно создавать тестовые классы с различными свойствами.
-Например, при фиксированных размерности задачи и числи локальных минимумов более сложный класс может быть сгенерирован путем сужения области притяжения точки глобального минимума или путем увеличения расстояния между этой точкой и вершиной параболоида.
-В проведенных экспериментах использовались значения $l=10$, $f^*=-1$, $\rho^*=0.2$ и $d^*=0.9$.
+By changing the specified parameters, you can create trial classes with different properties.
+For example, with a fixed dimensionality of the problem and the number of local minima, a more complex class can be generated by narrowing the region of attraction of the point of the global minimum or by increasing the distance between this point and the vertex of the paraboloid.
+In the experiments, the values $l=10$, $f^*=-1$, $\rho^*=0.2$ и $d^*=0.9$ were used.
 
-Для примера рассмотрим работу алгоритмов GSA и GSA-DT при решении одной из двумерных задач, порожденных генератором GKLS. 
-Линии уровня оптимизируемой функции, изображенные на figures \ref{fig3} и \ref{fig4}, наглядно показывают наличие десяти локальных экстремумов.
-При запуске алгоритмов использовались одинаковые параметры: $r=3.0$ из (\ref{R}) и $\epsilon = 10^{-2}$ в критерии остановки метода.
-На figures \ref{fig3} и \ref{fig4} точками черного цвета изображены точки search trials, выполненных методами в процессе решения задачи.
-При этом алгоритму GSA потребовалось 247 испытаний, тогда как алгоритму GSA-DT хватило 138 испытаний.
-Точкой красного цвета на рисунках отмечено точное решение задачи, а точкой желтого цвета - лучшее его приближение, найденное алгоритмом. 
-На рисунке \ref{fig4} точками зеленого цвета отмечены испытания, выполненные в рамках локального поиска. 
-Наглядно видно, что в методе с использованием деревьев решений для выделения областей притяжения локальных минимумов снимается проблема накопления точек испытаний в районе локальных экстремумов, присущая исходному алгоритму глобального поиска.
+As an example, consider the operation of the GSA and GSA-DT algorithms when solving one of the two-dimensional problems generated by the GKLS generator. 
+The level lines of the optimized function shown in Figures \ref{fig3} and \ref{fig4}, indicate the presence of ten local extrema. 
+When starting the algorithms, the same parameters were used:  $r=3.0$ from (\ref{R}) and $\epsilon = 10^{-2}$ in the stopping criterion of the method. 
+Black dots in Figures \ref{fig3} and \ref{fig4} show the points of the search trials performed by the methods in the process of solving the problem.
+In this case, the GSA algorithm required 247 trials, while the GSA-DT algorithm took 138 trials.
+The red dot in the figures marks the exact solution of the problem, and the yellow dot indicates the best approximation found by the algorithm. 
+Green dots in Figure \ref{fig4} indicate trials performed as part of a local search. 
+This graph demonstrates that using decision trees to identify areas of attraction of local minima removes the problem of accumulation of test points in the region of local extrema inherent in the original global search algorithm.
 
@@
 \includegraphics[width=0.6\linewidth]{GKLSAdaptiv6_30line.png}
-\caption{Решение задачи GKLS с помощью алгоритма GSA.}
+\caption{Using GSA algorithm to solve GKLS problems}
 \label{fig3}
@@
 \includegraphics[width=0.6\linewidth]{GKLSTree6_30line.png}
-\caption{Решение задачи GKLS с помощью алгоритма GSA-DT.}
+\caption{Using GSA-DT algorithm to solve GKLS problems}
 \label{fig4}
@@
 
-С помощью генератора GKLS было порождено по 300 тестовых задач размерностей $N=2,3,4$ (по 100 задач каждой размерности).
-Полученные серии задач были решены с помощью алгоритмов GSA и GSA-DT с параметром $r=5.0$ из (\ref{R}).
-Указанное значение параметра $r$ обеспечивает решение $100\%$ задач, при меньших значениях параметра некоторые задачи не были решены корректно.
-Отметим, что метод DIRECT
+We used GKLS to generate 300 test problems of dimensionalities $N=2,3,4$ (100 problems of each dimensionality).
+The resulting series of problems were solved using the GSA and GSA-DT algorithms with the parameter $r=5.0$ from (\ref{R}).
+The specified value of the parameter $r$ ensures the solution of  $100\%$ of the problems; at lower values of the parameter, some problems were not solved correctly.
+Note that the DIRECT method вот тут предложение обрывается внезапно
 
-В таблицах \ref{table:average_GKLS01} и \ref{table:average_GKLS002} приведено среднее число trials, потребовавшихся методам GSA и GSA-DT для корректного решения всех задач с точностью $\epsilon = 10^{-2}$ и $\epsilon = 2 \cdot 10^{-3}$ соответственно.
+Tables \ref{table:average_GKLS01} and \ref{table:average_GKLS002} show the average number of trials required by the GSA and GSA-DT methods to correctly solve all problems with an accuracy of $\epsilon = 10^{-2}$ и $\epsilon = 2 \cdot 10^{-3}$ respectively.
 
-Данные из таблиц подтверждают, что алгоритм глобального поиска, использующий выделение локальных экстремумов на основе машинного обучения, обеспечивает более быстрое решение многоэкстремальных задач, чем базовый алгоритм глобального поиска.
-При решении задач с грубой точностью ускорение составляет порядка $30\%$, тогда как при решении задач с высокой точностью процесс решения ускоряется от 2 до 6 раз.
+The data from the tables confirm that the global search algorithm based on the application of machine learning to extract local extremum provides a faster solution to multiextremum problems than the basic global search algorithm.
+For rough accuracy solutions, the acceleration is about  $30\%$, for high accuracy solutions, the process is accelerated from 2 to 6 times.
 
 \begin{specialtable}[H] 
-	\caption{Решение задач GKLS с точностью $\epsilon = 10^{-2}$}\label{table:average_GKLS01}
+	\caption{Solving GKLS problems with an accuracy  $\epsilon = 10^{-2}$}\label{table:average_GKLS01}
 	\center
@@
 \begin{specialtable}[H] 
-	\caption{Решение задач GKLS с точностью $\epsilon = 2 \cdot 10^{-3}$}\label{table:average_GKLS002}
+	\caption{Solving GKLS problems with an accuracy $\epsilon = 2 \cdot 10^{-3}$}\label{table:average_GKLS002}
 	\center
@@
 
-В статье рассмотрен эффективный детерминированный метод решения задач многоэкстремальной оптимизации -- the information-statistical global search algorithm. 
-Для данного алгоритма предложен новый способ ускорения его работы (в терминах числа итераций, требующихся для решения задачи с заданной точностью).
-Указанный способ основан на выявлении областей притяжения локальных минимумов целевой функции с использованием методов машинного обучения.
-Выявление областей притяжения и запуск в этих областях локального поиска позволяет существенным образом сократить число испытаний, требующееся методу для достижения глобальной сходимости.
-В рамках исследуемого подхода решение многомерных задач сводится к решению серии информационно-связанных одномерных подзадач, поэтому ключевым моментом является выявление локальных минимумов в одномерных задачах.
-Для этого используется аппроксимация целевой функции по точкам поисковых испытаний, построенная с помощью деревьев решений.
-Проведены вычислительные эксперименты на серии тестовых задач разной размерности с целью сравнения скорости работы исходного global search algorithm (GSA) и его модификации, которая использует деревья решений для выделения локальных минимумов целевой функции (GSA-DT).
-Результаты экспериментов показывают, что использование алгоритма GSA-DT может значительно (до 6 раз) сократить число испытаний, необходимое для решения задачи с заданной точностью.
+The article discusses an effective deterministic method for solving multiextremal optimization problems - the information-statistical global search algorithm.
+A new way of speeding up the operation of this algorithm was proposed (in terms of the number of iterations required to solve the problem with a given accuracy).
+This method is based on identifying areas of attraction of local minima of the objective function using machine learning methods.
+The identification of regions of attraction and the launch of local search in these regions can significantly reduce the number of trials required for the method to achieve global convergence.
+Within the framework of the investigated approach, solving multidimensional problems is reduced to solving a series of information-related one-dimensional subproblems; therefore, the key point is to identify local minima in one-dimensional problems.
+This is achieved by approximation of the objective function by points of search trials built using decision trees.
+Computational experiments were carried out on a series of test problems of different dimensions to compare the speed of the original global search algorithm (GSA) and its modification, which uses decision trees to identify local minima of the objective function (GSA-DT).
+The experimental results show that the use of the GSA-DT algorithm can significantly (up to 6 times) reduce the number of trials required to solve the problem with a given accuracy.
 
-Направлением дальнейших исследований будет использование более сложных моделей целевой фукнции для получения более точной ее аппроксимации.
-В качестве такого аппроксиматора планируется использовать artificial neural networks.
-Это потребует разработки новых методов выделения локальных экстремумов, т.к.
-аппроксимация функции с помощью нейронной сети является более сложной с этой точки зрения.
-Также планируется уделить внимание вопросам достоверности результатов, полученных с помощью методов машинного обучения.
-Для решенных модельных задач использование методов машинного обучения показывает хорошие результаты, однако вопрос о том, сохранится ли этот эффект в более сложных задачах, остается открытым.
+Further research into this issue will focus on using more complex models of the target function to obtain a more accurate approximation.
+We plan to use artificial neural networks as such an approximator.
+This will require the development of new methods for identifying local extrema since function approximation using a neural network is more complicated from this point of view.
+We also plan to pay attention to the issues of the reliability of the results obtained using machine learning methods.
+For the solved model problems, the use of machine learning methods shows good results, but the question of whether this effect will persist in more complex problems remains open.
 

### 2021-08-25 2021/Entropy/entropy.tex
fix some mistakes in translation

@@
 \abstract{
-The article features the study of global optimization problems and numerical methods of their solution.
-Such tasks are computationally expensive since the objective function can be multi-extremal, nondifferentiable, and, as a rule, given in the form of a "black box".
+The paper features the study of global optimization problems and numerical methods of their solution.
+Such problems are computationally expensive since the objective function can be multi-extremal, nondifferentiable, and, as a rule, given in the form of a ``black box''.
 The study used a deterministic algorithm for finding the global extremum.
 The said algorithm is based neither on the concept of multistart, nor nature-inspired algorithms.
-The article provides computational rules of the one-dimensional algorithm and the nested optimization scheme which could be applied for solving complex multidimensional problems.
+The article provides computational rules of the one-dimensional algorithm and the nested optimization scheme which could be applied for solving multidimensional problems.
 Note that the solution complexity of global optimization problems essentially depends on the presence of multiple local extrema.
-In this paper, we apply machine learning methods to identify areas of attraction of local minima.
-The use of local optimization algorithms in the selected areas can significantly accelerate the convergence of global search as it could reduce the number of search trials in the vicinity of local minima.
+In this paper, we apply machine learning methods to identify regions of attraction of local minima.
+The use of local optimization algorithms in the selected regions can significantly accelerate the convergence of global search as it could reduce the number of search trials in the vicinity of local minima.
 The results of computational experiments carried out on several hundred global optimization problems of different dimensionalities presented in the paper confirm the effect of accelerated convergence (in terms of the number of search trials required to solve a problem with a given accuracy).
@@
 Methods of machine learning were shown to be particularly effective for identifying the principal properties of the phenomena (for example, physical, economic, or social), which are stochastic by nature or contain some hidden parameters \cite{Golovenkin2020,Gonoskov2019}.
-ML is also successfully used to solve complex problems of computational mathematics, for example, for simulation of dynamical systems  \cite{Seleznev2019}, solution of ordinary, partial, or stochastic differential equations \cite{Lagaris1998,Blechschmidt2021,Xu2020} .
+ML is also successfully used to solve complex problems of computational mathematics, for example, for simulation of dynamical systems  \cite{Seleznev2019}, solution of ordinary, partial, or stochastic differential equations \cite{Lagaris1998,Blechschmidt2021,Xu2020}.
 
 For example, ML could be applied for solving such a complex problem of computational mathematics as global optimization.
-The solution to this class of problems, as a rule, cannnot be found analytically and, therefore, one needs to construct numerical methods to solve it.
+The solution to this class of problems, as a rule, cannot be found analytically and, therefore, one needs to construct numerical methods to solve it.
 
@@
 In many ways, they are related to the dimensionality and type of the objective function.
-Consequently, the most difficult tasks are those in which the objective function is multi-extremal, nondifferentiable, and, moreover, given in the form of a black box (i.e., in the form of some computational procedure, the input of which is an argument, and the output is the corresponding value of the function).
+Consequently, the most difficult problems are those in which the objective function is multi-extremal, nondifferentiable, and, moreover, given in the form of a ``black box'' (i.e., in the form of some computational procedure, the input of which is an argument, and the output is the corresponding value of the function).
 These complex problems are the main focus of this article.
@@
 Machine learning methods can be successfully applied to solve this problem.  
-For example, in \cite{RinnooyKan1987} methods of cluster analysis were used to select promising starting points
-In \cite{Cassioli2012} ], the area for launching the local method was allocated based on the classification of starting points using a support vector machine.
+For example, in \cite{RinnooyKan1987} methods of cluster analysis were used to select promising starting points.
+In \cite{Cassioli2012}, the area for launching the local method was allocated based on the classification of starting points using a support vector machine.
 
@@
 
-Note that the algorithms of the last two classes do not provide guaranteed convergence to the solution of the problem and are inferior to deterministic algorithms in terms of the quality of solution \cite{Kvasov2018,Sergeyev2018} ] (eg, measured by the number of correctly solved problems from a particular set).
+Note that the algorithms of the last two classes do not provide guaranteed convergence to the solution of the problem and are inferior to deterministic algorithms in terms of the quality of solution \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a particular set).
 Therefore, deterministic methods seem to be potentially more effective.
@@
 
-Several strategies have been proposed to speed up the operation of the global search algorithm (in terms of the number of iterations required to solve the problem with a given accuracy).
+Several strategies have been proposed to speedup the global search algorithm (in terms of the number of iterations required to solve the problem with a given accuracy).
 In this paper, we propose a new approach to acceleration based on identifying areas of attraction of local minima using machine learning methods.
@@
 \end{eqnarray}
-Problem (\ref{main_problem}) is considered under the assumption that the objective function is multi-extreme, is given in a form of a "black box" , and the calculation of its values is associated with solving the problem of numerical modeling, which makes the solution a labor-intensive operation.
+Problem (\ref{main_problem}) is considered under the assumption that the objective function is multi-extreme, is given in a form of a ``black box'', and the calculation of its values is associated with solving the problem of numerical simulation, which makes the solution a labor-intensive operation.
 
-A typical situation for many applied problems is when a limited change in the vector of parameters y causes a limited change in the values of $\varphi(y)$.
+A typical situation for many applied problems is when a limited change in the vector of parameters $y$ causes a limited change in the values of $\varphi(y)$.
 The mathematical model describing this premise is based on the assumption that the Lipschitz condition is satisfied
@@
 \]
-The common assumption for many approaches to the development of optimization algorithms is that the objective function is Lipschitzian  \cite{Jones1993,Pinter1996,Zilinskas2008,Evtushenko2009}.
+This assumption is typical for many approaches to the development of optimization algorithms \cite{Jones1993,Pinter1996,Zilinskas2008,Evtushenko2009}.
 At the same time, many known approaches are based on various methods of dividing the search domain into a system of subdomains and then choosing the most promising subdomain for placing the next trial (calculating the value of the objective function)  \cite{Jones2009,Zilinskas2010,Evtushenko2013,Kvasov2013,Paulavicius2016}.
@@
 As a result, when minimizing substantially multi-extremal functions, the numerical method must construct a coverage of the search domain.
-As a result, the number of nodes of this coverage increases exponentially with increasing dimensionality.
-This feature determines the high complexity of solving multiextremal optimization problems making dimensionality a critical factor that affects the complexity of their solution.
+The number of nodes of this coverage increases exponentially with increasing dimensionality.
+This feature determines the high complexity of solving multiextremal optimization problems making dimensionality a critical factor that affects the complexity of their solving.
 
-The dimensionality in multi-extreme optimization leads to many problems, so, scholars use a wide variety of approaches to reducing it.  
-For example, simplicial or diagonal partition of the search domain allows using methods for solving one-dimensional problems to solve the original multidimensional problem (see, for example,  \cite{PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
-Another well-known approach to dimensionality reduction is using the Peano space-filling curves to map the multidimensional domain onto a one-dimensional interval \cite{Strongin2000}.
+The dimensionality in multi-extremal optimization leads to many issues, so, scholars use a wide variety of approaches to reducing it.  
+For example, simplicial or diagonal partition of the search domain allows using methods for solving one-dimensional problems to solve the original multidimensional problem (see, for example,  \cite{PaulaviciusZilinskas2014,Sergeyev2017}).
+Another well-known approach to dimensionality reduction is using the Peano space-filling curves to map the multidimensional domain onto a one-dimensional interval \cite{Strongin2000,Sergeyev2013}.
 
 In this work, we will use another method based on the nested optimization scheme \cite{Shi2000,Grishagin2001,VanDam2010,Grishagin2015} and its generalization \cite{Grishagin2016,Grishagin2016_1}.
-The nested optimization scheme, on the one hand, does not worsen the properties of the minimized function (unlike reduction using Peano curves), and, on the other hand, does not require the use of complex data structures to support simplex or diagonal partitions of the feasible region
-At the same time, the nested optimization scheme makes it possible to reduce the solution of a multidimensional problem to the solution of a set of standard one-dimensional problems, which, then, can be solved by a wide range of one-dimensional global optimization algorithms.
+The nested optimization scheme, on the one hand, does not worsen the properties of the objective function (unlike reduction using Peano curves), and, on the other hand, does not require the use of complex data structures to support simplex or diagonal partitions of the feasible region.
+At the same time, the nested optimization scheme makes it possible to reduce the original multidimensional optimization problem to a family of recursively connected one-dimensional optimization subproblems, which can be solved by a wide range of one-dimensional global optimization algorithms.
 
@@
 \end{equation}
-with an objective function satisfying the Lipschitz condition
+with an objective function satisfying the Lipschitz condition.
 
 Here is the description of global search algorithm (GSA) for solving the basic problem in accordance with \cite{Strongin2000}.
-]. In the course of its work, GSA generates a sequence of points  $x^i$, at which the values of the objective function  $z^i=\varphi(x^i)$ are calculated. 
+In the course of its work, GSA generates a sequence of points  $x^i$, at which the values of the objective function  $z^i=\varphi(x^i)$ are calculated. 
 We will refer to the process of calculating the value of the objective function as \textit{trial}.
@@
 In accordance with the algorithm, the first two trials are carried out at the boundary points of the segment $[0,1]$, т.е. $x^0=0,\;x^1=1$. 
-At these points, the values of the objective function  $z^0=\varphi(x^0),\;z^1=\varphi(x^1)$ и уare calculated and the counter value is set to $k=1$. 
-The point of the next trial $x^{k+1}, k\geq 1,$ is selected in accordance with the following procedure:
+At these points, the values of the objective function  $z^0=\varphi(x^0),\;z^1=\varphi(x^1)$ are calculated and the counter value is set to $k=1$. 
+The point of the next trial $x^{k+1}, k\geq 1,$ is selected in accordance with the following procedure.
 
- Step 1. Reindex  (starting at 0) the points xl, $x^i,\:0\leq i\leq k$, of the trials conducted  in ascending order of the coordinate, i.e.
+ Step 1. Renumber  (starting at 0) the points $x^i,\:0\leq i\leq k$, of the trials conducted  in ascending order of the coordinate, i.e.
 \begin{equation}\label{xt}
@@
 \end{equation} 
-Compare the values of the objective function $z_i=\varphi(x_i), \; 0\leq i\leq k$ to the points $x_i, \; 0\leq i\leq k$, at which these values were calculated.
+Associate the values of the objective function $z_i=\varphi(x_i), \; 0\leq i\leq k,$ to the points $x_i, \; 0\leq i\leq k$, at which these values were calculated.
 
@@
 \end{equation}
-where $\Delta_i = x_i-x_{i-1}$. If the value calculated in accordance with this formula is equal to zero, then take $\mu = 1$.
+where $\Delta_i = x_i-x_{i-1}$. If the value calculated in accordance with (\ref{mu}) is equal to zero, then take $\mu = 1$.
 
@@
 
-Step 4. Find interval $(x_{t-1},x_t)$ with the maximum characteristics
+Step 4. Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
 \begin{equation}\label{MaxR}
@@
 \end{equation}
-If the maximum characteristics corresponds to several intervals, then choose the minimum number that satisfies as $t$ (\ref{MaxR}).
+If the maximum characteristic corresponds to several intervals, then choose the minimum number that satisfies (\ref{MaxR}) as $t$.
 
@@
 
-The algorithm stops when the condition  $\Delta_t<\epsilon$ is satisfied; here $t$ from (\ref{MaxR}), and $\epsilon>0$ is a given accuracy. 
-The following values are selected to evaluate the efficiency of the solution
+The algorithm stops when the condition  $\Delta_t<\epsilon$ is satisfied; here $t$ is from (\ref{MaxR}), and $\epsilon>0$ is a given accuracy. 
+The values
 \[
 z_k^\ast=\min_{0\leq i \leq k}\varphi(x^i), \ x_k^\ast=\arg \min_{0\leq i \leq
- k}\varphi(x^i).
+ k}\varphi(x^i)
 \] 
+are selected to estimate the solution
 
 The theoretical conditions that determine the convergence of the algorithm are presented in \cite{Strongin2000}. 
-The operation of the algorithm during the minimization of a specific multiextremal function, which is specified in accordance with formula  (\ref{hill}), is shown in Figure \ref{fig1}.
-The algorithm was launched with the parameter $r=2.2$ from (\ref{R}) and value $\epsilon = 10^{-3}$in the criterion for stopping the method.
+The work of the algorithm during the minimization of a specific multiextremal function, which is specified in accordance with formula  (\ref{hill}), is shown in Figure \ref{fig1}.
+The algorithm was launched with the parameter $r=2.2$ from (\ref{R}) and value $\epsilon = 10^{-3}$ in the method stopping condition.
 Figure \ref{fig1} shows the objective function graph and points of 71 search trials which GSA needed to solve the problem to the specified accuracy.
-This highlights the problem of all methods of global optimization - the concentration of points in the vicinity of local minima of the problem, which are not a global solution.
+This highlights the problem of all methods of global optimization --- the concentration of trial points in the vicinity of local minima of the problem, which are not a global solution.
 
@@
 
-\subsection{Machine Learning Regression as a tool for identifying areas of attraction of local extrema}\label{TreeGSA}
+\subsection{Machine Learning Regression as a tool for identifying attraction regions of local extrema}\label{TreeGSA}
 
@@
 A more powerful tool for this task is regression splines.
-When constructing a regression spline, the domain of definition is divided into K non-overlapping subdomains.
+When constructing a regression spline, the domain is divided into $K$ non-overlapping subdomains.
 In each of such domains, the function is approximated by a polynomial.
@@
 However, if there is a more complex relationship between them, then the decision tree can surpass the classical variants of regression in terms of the quality of the approximation.
-%Здесь нужен рисунок - функция Шекеля, и два варианта регрессии - полином и деревья решений. Шекель должен плохо полиномами представляться, т.к. там функции вида 1/x
 At the same time, the regression based on the decision tree makes it possible to easily identify the areas of attraction of local extrema with sufficient accuracy.
@@
 	\item Search domain $D$ is divided into $J$ non-overlapping subdomains  $D_1, D_2, ..., D_J$, provided that $D = \bigcup_{j=1}^{J}{D_j}$.
-	\item Any value falling into the subdomain $D_j$, i.e. $x \in D_j$, is matched to the average value  $c_j$ based on the training trial values that fall into this subdomain. 
+	\item Any value falling into the subdomain $D_j$, i.e. $x \in D_j$, is matched to the average value  $c_j$ based on the training trials that fall into this subdomain. 
 \end{itemize}
@@
 Next, the  $c_j$ values corresponding to neighboring subdomains are compared. If one of the conditions is met
-\begin{gather}\label{LocalPointC}
+\begin{gather}
 c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3} \leq c_{j+4}, \; j=1; \nonumber \\ 
 c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3}, \; j=2;  \nonumber \\  
-c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2}, \; 3 \leq j \leq J-2; \\ 
+c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2}, \; 3 \leq j \leq J-2; \label{LPC} \\ 
 c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1}, \; j=J-1; \nonumber \\ 
@@
 
-In order to modify the global search algorithm from subsection  \ref{CoreGSA} to exclude the regions of attraction of local minima, we associate each test point  $x^i$, obtained during the operation of the algorithm with an additional criterion $q^i \in \{0,1,2\}$, which will characterize the properties of this point.
+In order to modify the global search algorithm from subsection  \ref{CoreGSA} to exclude the regions of attraction of local minima, we associate each trial point  $x^i$ obtained during the operation of the algorithm with an additional attribute $q^i \in \{0,1,2\}$, which will characterize the properties of this point.
 The value  $q^i=0$ is assigned by default and indicates that point $x^i$ is obtained as a result of rule (\ref{xk1}) of the global search algorithm.
@@
 
-Steps 1-5 of the GSA-DT algorithm are the same as steps 1-5.
+Steps 1-5 of the GSA-DT algorithm are the same as steps 1-5 of GSA.
 
@@
 Step 7.
-Construct a decision tree based on the results of the tests performed, and obtain the corresponding piecewise linear approximation, which assigns the value $c_1, c_2, ..., c_J$ to each subdomain  $D_1, D_2, ..., D_J$.
+Construct a decision tree based on the results of the trials performed, and obtain the corresponding piecewise constant approximation, which assigns the value $c_1, c_2, ..., c_J$ to each subdomain  $D_1, D_2, ..., D_J$.
 
 Step 8.
-For the point $x^{k+1}$ of the current trial, find a number $j$ such that $x^{k+1} \in D_j$ and check whether the condition  (\ref{LocalPointC}) is satisfied.If condition (\ref{LocalPointC}) is satisfied, start a local search in the domain $D_j$ from the point  $x^{k+1}$.
+For the point $x^{k+1}$ of the current trial, find a number $j$ such that $x^{k+1} \in D_j$ and check whether the condition  (\ref{LPC}) is satisfied.
+If condition (\ref{LPC}) is satisfied, start a local search in the domain $D_j$ from the point  $x^{k+1}$.
 The results of all trials performed during the local search at the points $x^{k+2}, ...,x^{k+k_{local}}$ are stored in the information base of the algorithm and are used at subsequent iterations. 
@@
 \begin{itemize}
-	\item $|x_{t} - x_{t-1}|<\epsilon$, where $t$ from (\ref{MaxR}), and $\epsilon>0$ is a given accuracy;
-	\item $|x^{k+1} - x_{t-1}|<\epsilon$ and $q_{t-1} = 2$,
-	\item $|x^{k+1}-x_{t}|<\epsilon$ and $q_{t} = 2$.
+	\item $|x_{t} - x_{t-1}|<\epsilon$;
+	\item $|x^{k+1} - x_{t-1}|<\epsilon$ and $q_{t-1} = 2$;
+	\item $|x^{k+1}-x_{t}|<\epsilon$ and $q_{t} = 2$;
 \end{itemize}
+where $t$ is from (\ref{MaxR}), and $\epsilon>0$ is a given accuracy.
 
-For example, consider the operation of the GSA-DT algorithm with minimization of the same multi-extreme function, which is presented  Figure \ref{fig1}.
+For example, consider the work of the GSA-DT algorithm with minimization of the same multi-extremal function, which is presented in Figure \ref{fig1}.
 The same parameters were used at the start of the algorithm: the parameter  $r=2.2$ from (\ref{R}) and value $\epsilon = 10^{-3}$ in the stopping criterion of the method.
@@
 \end{equation}
-However, each calculation of the value of the function $y_1$ at some fixed point y1 presupposes the solution of the one-dimensional optimization problem of the second level
+However, each calculation of the value of the function $\varphi_1$ at some fixed point $y_1$ presupposes the solution of the one-dimensional optimization problem of the second level
 \begin{equation}
@@
 
-The solution of the set of subtasks arising in the nested optimization scheme  (\ref{nested_i}) can be organized in different ways. 
+The solution of the set of subproblems arising in the nested optimization scheme  (\ref{nested_i}) can be organized in different ways. 
 The obvious way (described in detail in \cite{Grishagin2001,Grishagin2015}) is based on solving subproblems in accordance with the recursive order of their generation.
@@
 The construction of this tree occurs dynamically in the process of solving the original problem (\ref{main_problem}).
-In this case, the calculation of one value of the function $\varphi_i(y_1,y_2,...,y_i)$ at the $i$-th ith level requires a complete solution of all problems of one of the subtrees of level  $i+1$.
+In this case, the calculation of one value of the function $\varphi_i(y_1,y_2,...,y_i)$ at the $i$-th level requires a complete solution of all problems of one of the subtrees of level  $i+1$.
 
@@
 the maximum characteristic from the characteristics of the intervals formed in this task.
-The higher the values of this characteristic, the more promising is the search for the global minimum of the original problem in the given subproblem (\ref{main_problem}).Therefore, at each iteration, a subproblem with the maximum characteristic is selected, and a trial is carried out in it.
-This either leads to the calculation of the value of the objective function  $\varphi(y)$(if the selected subtask belonged to the level  $j=M$), or generates new subproblems according to  (\ref{nested_i}) for $j\leq M-1$.
+The higher the values of this characteristic, the more promising is the search for the global minimum of the original problem (\ref{main_problem}) in the given subproblem.Therefore, at each iteration, a subproblem with the maximum characteristic is selected, and a trial is carried out in it.
+This either leads to the calculation of the value of the objective function  $\varphi(y)$ (if the selected subtask belonged to the level  $j=M$), or generates new subproblems according to (\ref{nested_i}) for $j\leq M-1$.
 In the latter case, new generated tasks are added to the current set of tasks, their characteristics are calculated, and the process is repeated.
@@
 a total of 16 CPU cores are available on the node).
-All the algorithms were implemented in Cd— +; GCC 5.5.0 was used for compilation on the supercomputer.
+All the algorithms were implemented in C++; GCC 5.5.0 was used for compilation on the supercomputer.
 
-The traditional approach to assessing the effectiveness of global optimization methods is based on using these methods to find the numerical solution of a series of a certain class of problems. 
+The traditional approach to assessing the effectiveness of global optimization methods is based on using these methods to find the numerical solution of a series of problems. 
 In this case, the assumption is that a certain algorithm is used to generate the next problem to be solved. 
 Typical examples of such test function classes are Shekel and Hill functions. 
-The first of them (denoted $F_{SH}$) is based on the formula.
+The first of them (denoted $F_{SH}$) is based on the formula
 \begin{equation}\label{shekel}
@@
 \begin{equation}\label{hill}
-  \varphi(x)=\sum_{j=1}^{14}(A_j\sin(2j\pi x) + B_j\cos(2j\pi x)),\: x\in[0,1]
+  \varphi(x)=\sum_{j=1}^{14}(A_j\sin(2j\pi x) + B_j\cos(2j\pi x)),\: x\in[0,1],
 \end{equation}
@@
 Let us compare the basic global search algorithm (GSA) and its decision tree-based modification (GSA-DT) with the well-known DIRECT global optimization algorithm  \cite{Jones2009}.  
-The decision trees in the GSA-DT algorithm were built using the OpenCV 4.5.1 library (class cv :: ml :: DTrees).
+The decision trees in the GSA-DT algorithm were built using the OpenCV 4.5.1 library (class cv::ml::DTrees).
 A regression function was constructed by a single decision tree.
 This allowed us to obtain a piecewise constant approximation of the objective function in which one or several trial points corresponded to each leaf node of the tree.
-The tree was built without any limitations on maximum depth (MaxDepth); the accuracy of tree construction (RegressionAccuracy) was 10-3; the minimum number of trial points in the tree node (MinSampleCount) was equal to one; all other parameters were set by default.
+The tree was built without any limitations on maximum depth (MaxDepth); the accuracy of tree construction (RegressionAccuracy) was $10^{-3}$; the minimum number of trial points in the tree node (MinSampleCount) was equal to one; all other parameters were set by default.
 Stopping tree building occurred if all absolute differences between an estimated value in a node and values of train samples in this node are less than accuracy.
@@
 The problem will be considered correctly solved if after stopping the method by accuracy (i.e.
-when the length of the current search interval becomes less than $\epsilon$) the current estimate of the optimum  $x_k^*$ lies in the $\epsilon$-neighborhood of the known solution of the problem  $x^*$, *, i.e.
+when the length of the current search interval becomes less than $\epsilon$) the current estimate of the optimum  $x_k^*$ lies in the $\epsilon$-neighborhood of the known solution of the problem  $x^*$, i.e.
 if the condition  $|x^*-x_k^*| \leq \epsilon$ is satisfied.
@@
 The procedure for generating test functions is based on using polynomials to redefine a convex quadratic function (paraboloid).
-Trial functions are defined by five parameters:
+Test functions are defined by five parameters:
 \begin{itemize}
 	\item dimensionality of the problem  $N$;
-	\item the number of local minima is $l$;
+	\item the number of local minima $l$;
 	\item value of the global minimum $f^*$;
 	\item radius of the area of attraction of the global optimizer $\rho^*$;
-	\item рthe distance between the global optimizer and the vertex of the paraboloid $d^*$.
+	\item the distance between the global optimizer and the vertex of the paraboloid $d^*$.
 \end{itemize}
-By changing the specified parameters, you can create trial classes with different properties.
+By changing the specified parameters, one can create test classes with different properties.
 For example, with a fixed dimensionality of the problem and the number of local minima, a more complex class can be generated by narrowing the region of attraction of the point of the global minimum or by increasing the distance between this point and the vertex of the paraboloid.
-In the experiments, the values $l=10$, $f^*=-1$, $\rho^*=0.2$ и $d^*=0.9$ were used.
+In the experiments, the values $l=10$, $f^*=-1$, $\rho^*=0.2$ and $d^*=0.9$ were used.
 
 As an example, consider the operation of the GSA and GSA-DT algorithms when solving one of the two-dimensional problems generated by the GKLS generator. 
-The level lines of the optimized function shown in Figures \ref{fig3} and \ref{fig4}, indicate the presence of ten local extrema. 
+The level lines of the objective function shown in Figures \ref{fig3} and \ref{fig4}, indicate the presence of ten local extrema. 
 When starting the algorithms, the same parameters were used:  $r=3.0$ from (\ref{R}) and $\epsilon = 10^{-2}$ in the stopping criterion of the method. 
@@
 The specified value of the parameter $r$ ensures the solution of  $100\%$ of the problems; at lower values of the parameter, some problems were not solved correctly.
-Note that the DIRECT method вот тут предложение обрывается внезапно
 
@@
 
-The article discusses an effective deterministic method for solving multiextremal optimization problems - the information-statistical global search algorithm.
-A new way of speeding up the operation of this algorithm was proposed (in terms of the number of iterations required to solve the problem with a given accuracy).
+The article discusses an efficient deterministic method for solving multiextremal optimization problems --- the information-statistical global search algorithm.
+A new way of speeding up the operation of this algorithm was proposed (in terms of the number of trials required to solve the problem with a given accuracy).
 This method is based on identifying areas of attraction of local minima of the objective function using machine learning methods.
@@
 Within the framework of the investigated approach, solving multidimensional problems is reduced to solving a series of information-related one-dimensional subproblems; therefore, the key point is to identify local minima in one-dimensional problems.
-This is achieved by approximation of the objective function by points of search trials built using decision trees.
+This is achieved by approximation of the objective function built using decision trees.
 Computational experiments were carried out on a series of test problems of different dimensions to compare the speed of the original global search algorithm (GSA) and its modification, which uses decision trees to identify local minima of the objective function (GSA-DT).
@@
 
-Further research into this issue will focus on using more complex models of the target function to obtain a more accurate approximation.
+Further research into this issue will focus on using more complex models of the objective function to obtain a more accurate approximation.
 We plan to use artificial neural networks as such an approximator.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Conceptualization and methodology, K.B.; software and validation, I.L.; formal analysis, K.B.; investigation, I.L.; data curation, I.L.; writing---original draft preparation, K.B.; writing---review and editing, K.B.; visualization, I.L.; funding acquisition, K.B.
+\authorcontributions{Conceptualization and methodology, K.B.; software and validation, I.L.; formal analysis, K.B.; investigation, I.L.; data curation, I.L.; writing --- original draft preparation, K.B.; writing- -- review and editing, K.B.; visualization, I.L.; funding acquisition, K.B.
 All authors have read and agreed to the published version of the manuscript.}

### 2021-08-26 2021/Entropy/entropy.tex
восстановил себя

@@
 \newcommand{\orcidauthorB}{0000-0002-8736-0652} % Add \orcidB{} behind the author's name
-%\newcommand{\orcidauthorC}{0000-0002-4013-2329} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorC}{0000-0001-6776-0096} % Add \orcidB{} behind the author's name
 
 % Authors, for the paper (add full first names)
-\Author{Konstantin Barkalov $^{1}$*\orcidA{} and Ilya Lebedev $^{1}$\orcidB{}}
+\Author{Konstantin Barkalov $^{1}$*\orcidA{} , Ilya Lebedev $^{1}$\orcidB{} and Evgeny Kozinov $^{1}$\orcidC{}}
 
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Konstantin Barkalov and Ilya Lebedev}
+\AuthorNames{Konstantin Barkalov, Ilya Lebedev and Evgeny Kozinov}
 
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Barkalov, K.; Lebedev, I.}
+\AuthorCitation{Barkalov, K.; Lebedev, I.; Kozinov, E.}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
@@
 \address[1]{%
-$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhny Novgorod, Russia; ilya.lebedev@itmm.unn.ru (I.L.)}
+$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhny Novgorod, Russia; ilya.lebedev@itmm.unn.ru (I.L.), evgeny.kozinov@itmm.unn.ru (E.A.)}
 %$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Conceptualization and methodology, K.B.; software and validation, I.L.; formal analysis, K.B.; investigation, I.L.; data curation, I.L.; writing --- original draft preparation, K.B.; writing- -- review and editing, K.B.; visualization, I.L.; funding acquisition, K.B.
+\authorcontributions{Conceptualization and methodology, K.B.; software and validation, I.L., E.A.; formal analysis, K.B.; investigation, I.L.; data curation, I.L.; writing --- original draft preparation, K.B.; writing- -- review and editing, K.B.; visualization, I.L.; funding acquisition, K.B.
 All authors have read and agreed to the published version of the manuscript.}

### 2021-08-27 2021/Entropy/entropy.tex
some fix
fix accuracy

@@
 \address[1]{%
-$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhny Novgorod, Russia; ilya.lebedev@itmm.unn.ru (I.L.), evgeny.kozinov@itmm.unn.ru (E.A.)}
+$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhny Novgorod, Russia; ilya.lebedev@itmm.unn.ru (I.L.), evgeny.kozinov@itmm.unn.ru (E.K.)}
 %$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
@@
 \begin{equation}\label{uni_problem}
-\varphi^\ast = \varphi(x^\ast)=\min{\left\{\varphi(x):x\in \left[0,1\right] 
+\varphi^\ast = \varphi(x^\ast)=\min{\left\{\varphi(x):x\in \left[a,b\right], a<b, a,b \in R
 \right\}}
@@
 
-In accordance with the algorithm, the first two trials are carried out at the boundary points of the segment $[0,1]$, т.е. $x^0=0,\;x^1=1$. 
+In accordance with the algorithm, the first two trials are carried out at the boundary points of the segment $[a,b]$, i.e. $x^0=a,\;x^1=b$. 
 At these points, the values of the objective function  $z^0=\varphi(x^0),\;z^1=\varphi(x^1)$ are calculated and the counter value is set to $k=1$. 
@@
 \begin{equation}\label{xt}
-0=x_0<x_1<\ldots <x_{k}=1.
+a=x_0<x_1<\ldots <x_{k}=b.
 \end{equation} 
@@
 The problem will be considered correctly solved if after stopping the method by accuracy (i.e.
-when the length of the current search interval becomes less than $\epsilon$) the current estimate of the optimum  $x_k^*$ lies in the $\epsilon$-neighborhood of the known solution of the problem  $x^*$, i.e.
-if the condition  $|x^*-x_k^*| \leq \epsilon$ is satisfied.
+when the length of the current search interval becomes less than $\epsilon \cdot \left|b-a\right| $) the current estimate of the optimum  $x_k^*$ lies in the $\epsilon$-neighborhood of the known solution of the problem  $x^*$, i.e.
+if the condition  $|x^*-x_k^*| \leq \epsilon \cdot \left|b-a\right|$ is satisfied.
 
@@
 At the same time, GSA-DT outperforms the DIRECT method both in the average number of search trials and in the number of correctly solved problems.
-In particular, if we use the accuracy $\epsilon = 10^{-2}$, then the DIRECT method stops too early and does not find a global solution to many problems.
+In particular, if we use the accuracy $\epsilon = 10^{-2}\left|b-a\right|$, then the DIRECT method stops too early and does not find a global solution to many problems.
 Therefore, in further experiments in which multidimensional problems are solved, we will not use DIRECT for comparison, since when solving multidimensional problems with stopping by accuracy, this method will provide a correct solution to no more than  $50\%$ of problems.
@@
 The level lines of the objective function shown in Figures \ref{fig3} and \ref{fig4}, indicate the presence of ten local extrema. 
-When starting the algorithms, the same parameters were used:  $r=3.0$ from (\ref{R}) and $\epsilon = 10^{-2}$ in the stopping criterion of the method. 
+When starting the algorithms, the same parameters were used:  $r=3.0$ from (\ref{R}) and $\epsilon = 10^{-2}\left|b-a\right|$ in the stopping criterion of the method. 
 Black dots in Figures \ref{fig3} and \ref{fig4} show the points of the search trials performed by the methods in the process of solving the problem.
@@
 
-Tables \ref{table:average_GKLS01} and \ref{table:average_GKLS002} show the average number of trials required by the GSA and GSA-DT methods to correctly solve all problems with an accuracy of $\epsilon = 10^{-2}$ и $\epsilon = 2 \cdot 10^{-3}$ respectively.
+Tables \ref{table:average_GKLS01} and \ref{table:average_GKLS002} show the average number of trials required by the GSA and GSA-DT methods to correctly solve all problems with an accuracy of $\epsilon = 10^{-2}\left|b-a\right|$ и $\epsilon = 2 \cdot 10^{-3}\left|b-a\right|$ respectively.
 
@@
 \begin{specialtable}[H] 
-	\caption{Solving GKLS problems with an accuracy  $\epsilon = 10^{-2}$}\label{table:average_GKLS01}
+	\caption{Solving GKLS problems with an accuracy  $\epsilon = 10^{-2}\left|b-a\right|$}\label{table:average_GKLS01}
 	\center
@@
 \begin{specialtable}[H] 
-	\caption{Solving GKLS problems with an accuracy $\epsilon = 2 \cdot 10^{-3}$}\label{table:average_GKLS002}
+	\caption{Solving GKLS problems with an accuracy $\epsilon = 2 \cdot 10^{-3}\left|b-a\right|$}\label{table:average_GKLS002}
 	\center
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Conceptualization and methodology, K.B.; software and validation, I.L., E.A.; formal analysis, K.B.; investigation, I.L.; data curation, I.L.; writing --- original draft preparation, K.B.; writing- -- review and editing, K.B.; visualization, I.L.; funding acquisition, K.B.
+\authorcontributions{Conceptualization and methodology, K.B.; software and validation, I.L., E.K.; formal analysis, K.B.; investigation, I.L.; data curation, I.L.; writing --- original draft preparation, K.B.; writing- -- review and editing, K.B.; visualization, I.L.; funding acquisition, K.B.
 All authors have read and agreed to the published version of the manuscript.}

### 2021-08-28 2021/Entropy/entropy.tex
before submission

@@
 %%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
-\usepackage{inputenc}
-\usepackage[T2A,T1]{fontenc}
-\usepackage[english,russian]{babel}
-\usepackage{cmap}
+%\usepackage{inputenc}
+%\usepackage[T2A,T1]{fontenc}
+%\usepackage[english,russian]{babel}
+%\usepackage{cmap}
 %%%%
@@
 
-For example, ML could be applied for solving such a complex problem of computational mathematics as global optimization.
+Particularly, ML could be applied for solving such a complex problem of computational mathematics as global optimization.
 The solution to this class of problems, as a rule, cannot be found analytically and, therefore, one needs to construct numerical methods to solve it.
@@
 For example, in \cite{RinnooyKan1987} methods of cluster analysis were used to select promising starting points.
-In \cite{Cassioli2012}, the area for launching the local method was allocated based on the classification of starting points using a support vector machine.
+In \cite{Cassioli2012} the area for starting the local method was allocated based on the classification of starting points using a support vector machine.
 
@@
 One of the efficient deterministic methods for solving multiextremal optimization problems is \textit{the information-statistical global search algorithm} \cite{Strongin2000}.
-This method initially proposed for solving unconstrained optimization problems was successfully generalized to the classes of optimization problems with non-convex constraints and multicriteria optimization problems.
+This method initially proposed for solving unconstrained optimization problems was successfully generalized to the classes of optimization problems with non-convex constraints \cite{Barkalov2002} and multicriteria optimization problems \cite{Gergel2020}.
 For different versions of the algorithm, parallelization methods taking into account the architecture of modern computing systems were also suggested \cite{Barkalov2016,globalizerSystem,Strongin2018}.
@@
 \end{eqnarray}
-Problem (\ref{main_problem}) is considered under the assumption that the objective function is multi-extreme, is given in a form of a ``black box'', and the calculation of its values is associated with solving the problem of numerical simulation, which makes the solution a labor-intensive operation.
+Problem (\ref{main_problem}) is considered under the assumption that the objective function is multi-extremal, is given in a form of a ``black box'', and the calculation of its values is associated with solving the problem of numerical simulation, which makes the solution a labor-intensive operation.
 
@@
 
-Tables \ref{table:average_GKLS01} and \ref{table:average_GKLS002} show the average number of trials required by the GSA and GSA-DT methods to correctly solve all problems with an accuracy of $\epsilon = 10^{-2}\left|b-a\right|$ и $\epsilon = 2 \cdot 10^{-3}\left|b-a\right|$ respectively.
+Tables \ref{table:average_GKLS01} and \ref{table:average_GKLS002} show the average number of trials required by the GSA and GSA-DT methods to correctly solve all problems with an accuracy of $\epsilon = 10^{-2}\left|b-a\right|$ and $\epsilon = 2 \cdot 10^{-3}\left|b-a\right|$ respectively.
 

### 2021-09-13 2021/ICCS_BarkalovUsova/iccs.tex
fix one wrong symbol
fix grant number

--- 2021-06-15 2021/ICCS_BarkalovUsova/iccs.tex
@@
 %
-\title{A search algorithm for the global extremum \break of a discontinuous function\thanks{This study was supported by the Russian Science Foundation, project No.\,21-11-00204.}}
+\title{A search algorithm for the global extremum \break of a discontinuous function\thanks{This work was supported by the Russian Science Foundation, project No.\,21-11-00204.}}
 %
-\titlerunning{An algorithm for searching the global minimum}
+\titlerunning{A search algorithm for the global extremum}
 % If the paper title is too long for the running head, you can set 

### 2021-09-15 2021/Entropy/entropy.tex
rewrite some part of the Introduction

@@
 %%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
-%\usepackage{inputenc}
-%\usepackage[T2A,T1]{fontenc}
-%\usepackage[english,russian]{babel}
-%\usepackage{cmap}
+\usepackage{inputenc}
+\usepackage[T2A,T1]{fontenc}
+\usepackage[english,russian]{babel}
+\usepackage{cmap}
 %%%%
@@
 There are several approaches to the construction of numerical methods for solving global optimization problems. 
-%Надо ли добавить алгоритмы на основе метамоделей ?
 Some algorithms are based on the idea of a multistart: launching a local search either from different starting points or with varying parameters.
@@
 
+%Russian
+Методы машинного обучения активно используются в сочетании с алгоритмами bayesian optimization, основанными на использовании probabilistic surrogate models of the objective function.
+Детальный обзор данного направления развития методов глобальной оптмизации представлен в \cite{Archetti2019,Zhigljavsky2021}.
+
 Another popular class of methods for solving global optimization problems is metaheuristic algorithms. 
@@
 
-Note that the algorithms of the last two classes do not provide guaranteed convergence to the solution of the problem and are inferior to deterministic algorithms in terms of the quality of solution \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a particular set).
+Note that the algorithms of the last class do not provide guaranteed convergence to the solution of the problem and are inferior to deterministic algorithms in terms of the quality of solution \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a particular set).
 Therefore, deterministic methods seem to be potentially more effective.
 
-One of the efficient deterministic methods for solving multiextremal optimization problems is \textit{the information-statistical global search algorithm} \cite{Strongin2000}.
-This method initially proposed for solving unconstrained optimization problems was successfully generalized to the classes of optimization problems with non-convex constraints \cite{Barkalov2002} and multicriteria optimization problems \cite{Gergel2020}.
-For different versions of the algorithm, parallelization methods taking into account the architecture of modern computing systems were also suggested \cite{Barkalov2016,globalizerSystem,Strongin2018}.
+%Russian
+Настоящая работа направлена на дальнейшее развитие эффективного детерминированного метода глобальной оптимизации, известного как \textit{the information-statistical global search algorithm} \cite{Strongin2000}.
+Указанная здесь книга содержит результаты теоретических исследований метода, имеющих непосредственное значение для его практической реализации.
+В частности, в ней детально рассмотрены вопросы сходимости, выбора параметров и условия остановки алгоритма, и т.д.
+Отметим, что алгоритм глобального поиска первоначально был ориентирован на решение unconstrained optimization problems.
+Впоследствии он был обобщен для решения problems with non-convex constraints\cite{Barkalov2002} and multicriteria optimization problems \cite{Gergel2020}.
+Одновременно с этим были предложены различные параллельные варианты указанных алгоритмов, которые могут быть использованы на современных суперкомпьютерах \cite{Barkalov2016,globalizerSystem,Strongin2018}.
 
@@
 \]
-This assumption is typical for many approaches to the development of optimization algorithms \cite{Jones1993,Pinter1996,Zilinskas2008,Evtushenko2009}.
-At the same time, many known approaches are based on various methods of dividing the search domain into a system of subdomains and then choosing the most promising subdomain for placing the next trial (calculating the value of the objective function)  \cite{Jones2009,Zilinskas2010,Evtushenko2013,Kvasov2013,Paulavicius2016}.
+This assumption is typical for many approaches to the development of optimization algorithms \cite{Jones1993,Pinter1996,Zilinskas2008,Evtushenko2009,Sergeyev2020}.
+At the same time, many known approaches are based on various methods of dividing the search domain into a system of subdomains and then choosing the most promising subdomain for placing the next trial (calculating the value of the objective function)  \cite{Jones2009,Zilinskas2010,Evtushenko2013,Kvasov2013,Paulavicius2016,Paulavicius2020}.
 An important property of global optimization problems is the fact that, in contrast to the problems of finding a local extremum, the global minimum is an integral characteristic of the problem being solved.

### 2021-09-16 2021/Entropy/entropy.tex
comment about stopping condition

@@
 
-Recall that the superscript corresponds to the number of the iteration at which the trial was carried out at this point, and the subscript corresponds to the number of the point in the row  (\ref{xt}).
+%Russian
+Использование additional attribute $q^i, \; 0 \leq i \leq k ,$ позволяет различать точки, полученные во время глобального ($q^i=0$) и во время локального ($q^i=1,2$) поиска.
+Интервалы, граничные точки которых имеют значение $q^i=1$ или $q^i=2$, будут пропускаться при глобальном поиске, т.е.
+в них не будут проводиться дальнейшие испытания.
+А выделенное значение $q^i=2$, соответствующее локальному минимуму, будет позволять использовать данные точки при проверке критерия остановки.
+Если по правилам глобального поиска новая точка испытания должна быть поставлена в заданной $\epsilon$-окрестности одного из найденных локальных минимумов, это будет соответствовать окончанию поиска.
 
 Let us now describe a modified global search algorithm that uses decision trees to isolate and exclude areas of attraction of local minima; we will further refer to this algorithm as GSA-DT.
+Recall that the superscript corresponds to the number of the iteration at which the trial was carried out at this point, and the subscript corresponds to the number of the point in the row  (\ref{xt}).
 
@@
 where $t$ is from (\ref{MaxR}), and $\epsilon>0$ is a given accuracy.
+
+%Russian
+Отметим, что указанный критерий остановки проверяется после Step 6, т.е.
+после вычисления по правилам глобального поиска точки $x^{k+1}$ очередного испытания, но до проведения самого испытания в ней.
+Смысл данного критерия заключается в следующем.
+Поиск останавиливается, когда интервал $(x_{t-1},x_t)$, в который попадает точка $x^{k+1}$, становится достаточно малым, либо когда точка $x^{k+1}$ попадает в малую окрестность одного из найденных локальных минимумов, который в этом случае будет одновременно являться и глобальным.
 
@@
 The number of unsolved problems is indicated in parentheses.
+%Russian
+В этих и последующих таблицах для метода GSA-DT приведено полное число испытаний, которые были выполнены как во время глобального, так и во время локального поиска.
+%Ebglish
 The experimental results show that with a rough solution to the problem, all methods show similar results in terms of the number of trials, while with a high solution accuracy, the GSA-DT algorithm requires 2 times fewer trials than its prototype.
@@
 
+%Russian
+Отметим, что DIRECT является одним из наиболее известных и популярных детерминированных методов решения задач глобальной оптмизации with ``black box'' objective function, и при достаточно большом числе поисковых испытаний DIRECT гарантированно найдет глобальное решение задачи.
+Однако если использовать критерий основки по точности, то метод может быстро остановиться в одном из локальных минимумов.
+Сказанное подтверждается приведенными в данном разделе результатами экспериментов.
+
+Если же говорить про детерминированные оптимизационные алгоритмы, реализованные in popular Computer Algebra Systems (CAS), то их сравнение с такими методами, как DIRECT или GSA, будет некорректным по следующей причине.
+
+Оптимизационные алгоритмы из CAS ориентированы на решение задач с целевой функцией, заданной явно, в виде формулы.
+Формульное задание целевой функции предполагает известным также и ее производные, что позволяет использовать методы первого порядка, обладающие значительно более быстрой сходимостью, чем методы нулевого порядка, к которым отноится и DIRECT, и GSA.
+Однако методы, требующие формульного задания целевой функции, не могут применять к решению большого класса прикладных задач, в которых вид целевой функции не известен, а ее значения вычисляются в результате численного моделирования.
+
 \begin{specialtable}[H] 
-	\caption{The average number of tests when minimizing Shekel test functions (the number of unsolved problems is indicated in parentheses))}\label{table:average_Shekel}
+	\caption{The average number of tests when minimizing Shekel test functions (the number of unsolved problems is indicated in parentheses)}\label{table:average_Shekel}
 	\center
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
+\section{Conclusions and Future Work}
 
@@
 For the solved model problems, the use of machine learning methods shows good results, but the question of whether this effect will persist in more complex problems remains open.
+
+%Russian
+Еще одним направлением дальнейшей работы будет являться комбинирование предложенного подхода к использованию локального поиска с традиционными способами ускорения алгоритмов глобальной оптмизации.
+Например, широко известным способом ускорения алгоритмов данного класса является использование local adaptive estimates of the Lipschitz constant $L_i$ в различных зонах $D_i \in D$ области поиска $D$ вместо одной глобальной оценки константы $L$ во всей области $D$ ( see \cite{Kvasov2020,Sergeyev2021}).
+Это позволяет алгоритму настраиваться на локальное поведение целевой функции в различных частях допустимой области и сокращает число поисковых испытаний, требующихся для достижения сходимости.
+Использование только глобальной информации о поведении целевой функции в ходе ее минимизации может значительно замедлить сходимость алгоритма к точке
+глобального минимума.
+В данном направлении также планируется использовать методы машинного обучения для выделения подобластей $D_i \in D$ и построения локальных оценок константы Липшица $L_i$ в выделенных подобластях.
 

### 2021-09-19 2021/Entropy/entropy.tex
answers for the comments in russian

@@
 \end{itemize}
-
 In fact, decision trees build a model of a function of the form
@@
 
-This model is, on the one hand, rather simple, and on the other hand, it adequately reflects the properties of the function under investigation (the presence or absence of local minima).
+%Russian
+В целом, можно сказать, что дерево принятия решений -- это бинарное дерево, в листьях которого находятся значения функции, а в остальных узлах -- условия перехода.
+В нашем случае при применении дерева решений для построения регрессии каждому узлу соответствует несколько результатов испытаний и значение узла $c_j$, которое вычисляется как среднее квадратичное значений в точках испытаний, приписанных к этому узлу.
+Итоговая кусочно-постоянная аппроксимация строится с использованием значений $c_j$, которые находятся в листьях дерева.
+
+Дерево строится рекурсивно, начиная с корневого узла. 
+В каждом узле применяется правило принятия решения: все данные  разбиваются на две группы (согласно некоторому правилу разбиения) и отправляются в левый и правый узлы-потомки.
+Затем процедура рекурсивно разделяет левый и правый узлы.
+Остановка рекурсии в узле происходит в одном из следующих случаев:
+\begin{itemize}
+	\item Количество приписанных к узлу точе испытаний становится меньше указанного порогового значения (мы использовали 1).
+	\item Сумма квадратов отклонений значений функции от значения $c_j$, приписанного к этому узлу, становится меньше заданной точности (мы использовали 0.0001).
+\end{itemize}
+Подробнее об алгоритме построения регрессии с использованием деревьев решений см., например, \cite{Breiman1984}.
+
+Регрессия, построенная с помощью деревьев решений,
+is, on the one hand, rather simple, and on the other hand, it adequately reflects the properties of the function under investigation (the presence or absence of local minima).
 The selection of regions of attraction of local minima using model (\ref{dtree}) can be organized as follows.
@@
 
-A brief description of the main steps of the adaptive dimensionality reduction scheme is as follows.
-
-Let nested subproblems of the form (\ref{nested_i}) be solved using the global search algorithm described in subsection  \ref{CoreGSA}.
+%A brief description of the main steps of the adaptive dimensionality reduction scheme is as follows.
+%Let nested subproblems of the form (\ref{nested_i}) be solved using the global search algorithm described in subsection  \ref{CoreGSA}.
 Then each subproblem  (\ref{nested_i})  can be assigned some numerical value called a characteristic of this problem.
@@
 Let us compare the basic global search algorithm (GSA) and its decision tree-based modification (GSA-DT) with the well-known DIRECT global optimization algorithm  \cite{Jones2009}.
+%Russian
+Выбор именного этого метода для сравнения объясняется следующим.
+DIRECT является одним из наиболее известных и популярных детерминированных методов решения задач глобальной оптмизации with ``black box'' objective function.
+Обзор различных модификаций метода, а также примеров решения задач приведен в \cite{Jones2021}.
+Для этого метода известно, что при достаточно большом числе поисковых испытаний DIRECT гарантированно найдет глобальное решение задачи.  
+Однако если использовать критерий основки по точности, то метод может быстро остановиться в одном из локальных минимумов.
+Сказанное подтверждается приведенными в данном разделе результатами экспериментов.
+
+Если же говорить про детерминированные оптимизационные алгоритмы, реализованные in popular Computer Algebra Systems (CAS), то их сравнение с такими методами, как DIRECT или GSA, будет некорректным по следующей причине.
+Оптимизационные алгоритмы из CAS ориентированы на решение задач с целевой функцией, заданной явно, в виде формулы.
+Формульное задание целевой функции предполагает известным также и ее производные, что позволяет использовать методы первого порядка, обладающие значительно более быстрой сходимостью, чем методы нулевого порядка, к которым отноится и DIRECT, и GSA.
+Но при этом методы первого порядка не гарантируют нахождение глобального минимума.
+Например, система Mathematica предлагает для решения задач глобальной оптмизации методы, которые гарантированно найдут глобальное решение задачи только в случае, если целевая функция и ограничения являются линейными или выпуклыми.
+Otherwise, the result may sometimes only be a local minimum.
+
+Более того, методы, требующие формульного задания целевой функции, не могут применяться к решению большого класса прикладных задач, в которых вид целевой функции не известен, а ее значения вычисляются в результате численного моделирования.
+Использование для решения задач подобного вида эвристических методов (таких как Differential Evolution, Simulated Annealing or Random Search) также не всегда приводит к решению задачи.
+These methods often do find the global minimum, but are not guaranteed to do so. В терминах числа корректно решенных задач эвристические методы проигрывают детерминированным \cite{Sergeyev2018}.
+
 The decision trees in the GSA-DT algorithm were built using the OpenCV 4.5.1 library (class cv::ml::DTrees).
@@
 В этих и последующих таблицах для метода GSA-DT приведено полное число испытаний, которые были выполнены как во время глобального, так и во время локального поиска.
+При решении задач с точностью $10^{-2}$, $10^{-3}$, $10^{-4}$ число запусков локального поиска при решении одной задачи составляло в среднем 2, 3 и 4 соответственно; а отношение числа испытаний, выполненных по правилам глобального поиска, к числу испытаний, выполненных локальным методом, составляло (при соответствующей точности) 7.2, 6.0, 4.5 для задач класса $F_{HL}$ и 3.2, 2.3, 1.9 для задач класса $F_{SH}$.
+
 %Ebglish
@@
 Therefore, in further experiments in which multidimensional problems are solved, we will not use DIRECT for comparison, since when solving multidimensional problems with stopping by accuracy, this method will provide a correct solution to no more than  $50\%$ of problems.
-
-%Russian
-Отметим, что DIRECT является одним из наиболее известных и популярных детерминированных методов решения задач глобальной оптмизации with ``black box'' objective function, и при достаточно большом числе поисковых испытаний DIRECT гарантированно найдет глобальное решение задачи.
-Однако если использовать критерий основки по точности, то метод может быстро остановиться в одном из локальных минимумов.
-Сказанное подтверждается приведенными в данном разделе результатами экспериментов.
-
-Если же говорить про детерминированные оптимизационные алгоритмы, реализованные in popular Computer Algebra Systems (CAS), то их сравнение с такими методами, как DIRECT или GSA, будет некорректным по следующей причине.
-
-Оптимизационные алгоритмы из CAS ориентированы на решение задач с целевой функцией, заданной явно, в виде формулы.
-Формульное задание целевой функции предполагает известным также и ее производные, что позволяет использовать методы первого порядка, обладающие значительно более быстрой сходимостью, чем методы нулевого порядка, к которым отноится и DIRECT, и GSA.
-Однако методы, требующие формульного задания целевой функции, не могут применять к решению большого класса прикладных задач, в которых вид целевой функции не известен, а ее значения вычисляются в результате численного моделирования.
 

### 2021-09-19 2021/Entropy/entropy.tex
Поправлена опечатка

@@
 Методы машинного обучения активно используются в сочетании с алгоритмами bayesian optimization, основанными на использовании probabilistic surrogate models of the objective function.
-Детальный обзор данного направления развития методов глобальной оптмизации представлен в \cite{Archetti2019,Zhigljavsky2021}.
+Детальный обзор данного направления развития методов глобальной оптимизации представлен в \cite{Archetti2019,Zhigljavsky2021}.
 
@@
 \begin{itemize}
-	\item Количество приписанных к узлу точе испытаний становится меньше указанного порогового значения (мы использовали 1).
-	\item Сумма квадратов отклонений значений функции от значения $c_j$, приписанного к этому узлу, становится меньше заданной точности (мы использовали 0.0001).
+	\item Количество приписанных к узлу точек испытаний становится меньше указанного порогового значения (мы использовали 1).
+	\item Сумма квадратов отклонений значений функции от значения $c_j$, приписанного к этому узлу, становится меньше заданной точности (мы использовали 0.001).
 \end{itemize}
@@
 Смысл данного критерия заключается в следующем.
-Поиск останавиливается, когда интервал $(x_{t-1},x_t)$, в который попадает точка $x^{k+1}$, становится достаточно малым, либо когда точка $x^{k+1}$ попадает в малую окрестность одного из найденных локальных минимумов, который в этом случае будет одновременно являться и глобальным.
+Поиск останавливается, когда интервал $(x_{t-1},x_t)$, в который попадает точка $x^{k+1}$, становится достаточно малым, либо когда точка $x^{k+1}$ попадает в малую окрестность одного из найденных локальных минимумов, который в этом случае будет одновременно являться и глобальным.
 
@@
 %Russian
-Выбор именного этого метода для сравнения объясняется следующим.
-DIRECT является одним из наиболее известных и популярных детерминированных методов решения задач глобальной оптмизации with ``black box'' objective function.
+Выбор именно этого метода для сравнения объясняется следующим.
+DIRECT является одним из наиболее известных и популярных детерминированных методов решения задач глобальной оптимизации with ``black box'' objective function.
 Обзор различных модификаций метода, а также примеров решения задач приведен в \cite{Jones2021}.
 Для этого метода известно, что при достаточно большом числе поисковых испытаний DIRECT гарантированно найдет глобальное решение задачи.  
-Однако если использовать критерий основки по точности, то метод может быстро остановиться в одном из локальных минимумов.
+Однако если использовать критерий остановки по точности, то метод может быстро остановиться в одном из локальных минимумов.
 Сказанное подтверждается приведенными в данном разделе результатами экспериментов.
@@
 Оптимизационные алгоритмы из CAS ориентированы на решение задач с целевой функцией, заданной явно, в виде формулы.
-Формульное задание целевой функции предполагает известным также и ее производные, что позволяет использовать методы первого порядка, обладающие значительно более быстрой сходимостью, чем методы нулевого порядка, к которым отноится и DIRECT, и GSA.
+Формульное задание целевой функции предполагает известным также и ее производные, что позволяет использовать методы первого порядка, обладающие значительно более быстрой сходимостью, чем методы нулевого порядка, к которым относится и DIRECT, и GSA.
 Но при этом методы первого порядка не гарантируют нахождение глобального минимума.
@@
 %Russian
-Еще одним направлением дальнейшей работы будет являться комбинирование предложенного подхода к использованию локального поиска с традиционными способами ускорения алгоритмов глобальной оптмизации.
+Еще одним направлением дальнейшей работы будет являться комбинирование предложенного подхода к использованию локального поиска с традиционными способами ускорения алгоритмов глобальной оптимизации.
 Например, широко известным способом ускорения алгоритмов данного класса является использование local adaptive estimates of the Lipschitz constant $L_i$ в различных зонах $D_i \in D$ области поиска $D$ вместо одной глобальной оценки константы $L$ во всей области $D$ ( see \cite{Kvasov2020,Sergeyev2021}).

### 2021-09-24 2021/ICCS_BarkalovUsova/iccs.tex
answers in Russian

@@
 %%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
-%\usepackage{inputenc}
-%\usepackage[english,russian]{babel}
-%\usepackage{cmap}
+\usepackage{inputenc}
+\usepackage[english,russian]{babel}
+\usepackage{cmap}
 %%%%
@@
 
-Global optimization problems involving many local extrema arise in many areas of applications (see, e.g., \cite{Cavoretto2021,Kvasov2015,Kvasov2013,Modorskii2016}).
+Global optimization problems involving many local extrema arise in many areas of applications (see, e.g., \cite{Cavoretto2021,Kvasov2013,Kvasov2015,Modorskii2016}).
 As a rule, the calculation of even a single value of the objective function in such problems requires considerable time, since the analytical form of the function is not known and the calculation of its values involves time-consuming numerical modelling.
@@
 Firstly, as has already been mentioned, they arise in many applications (see, e.g., \cite{Gillard2017,Sergeyev2020}).
-Secondly, many approaches to solving multidimensional optimization problems are, in one way or another, based on reducing the solution of the original multidimensional problem to solving a series of related one-dimensional optimization problems (see, for example, popular approaches in \cite{Pinter1996,PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).
+Secondly, many approaches to solving multidimensional optimization problems are, in one way or another, based on reducing the solution of the original multidimensional problem to solving a series of related one-dimensional optimization problems (see, for example, popular approaches in \cite{PaulaviciusZilinskas2014,Pinter1996,Sergeyev2017,Sergeyev2013}).
 
@@
 \item For each interval $(x_{i-1},x_i), \; 1\leq i \leq m,$ calculate the characteristic
-%\begin{equation}\label{R_i}
-\[
+\begin{equation}\label{R_i}
 \begin{gathered}
@@
 \end{gathered}
-%\end{equation}
-\]
+\end{equation}
 where $r>1$ is the method parameter.
@@
 
-Note that the erroneous assignment of some discontinuity points from the series (\ref{points}) will not lead to the loss of convergence of the method.
+%Russian
+
+\textbf{Remark 1}.
+Еще раз отметим, что в соответствии с (\ref{gamma}) значение $\gamma_i+\gamma_{i-1}$ может быть равно либо 0 (обе граничных точки интервала принадлежат области непрерывности функции), либо 1 (одна из граничных точек является точкой разрыва).
+В этом случае формула (\ref{R_i}) иметь вид
+\[
+R(i)=\Delta_i+\frac{ (z_i-z_{i-1} )^2}{(r\mu)^2 \Delta_i }-\frac{2(z_i + z_{i-1})}{r\mu},
+\]
+в случае отстутсвия точке разрыва и 
+\[
+R(i)=2\Delta_i - \frac{4 z_i }{r\mu}, \; R(i)=2\Delta_i - \frac{4 z_{i-1} }{r\mu},
+\]
+в случае разрыва в точке $x_{i-1}$ или $x_{i}$ соответственно.
+
+\textbf{Remark 2}.
+Приведенные выше формулы для вычисления характеристик интервалов могут быть использованы (в модифицированной форме) в алгоритме решения задач невыпуклыми ограничениями \cite{Sergeyev1995,Sergeyev2006_1,Sergeyev2007}.
+Здесь решение задачи с частично определенными ограничениями неявно сводится к минимизации разрывной функции, в которой точками разрыва являются границы допустимого множества решений задачи.
+
+\textbf{Remark 3}.
+The erroneous assignment of some discontinuity points from the series (\ref{points}) will not lead to the loss of convergence of the method.
 The algorithm will interpret such points as a recoverable discontinuity case since the left and right limits of the objective function at such points coincide.
@@
 \end{equation}
-where $Q>1$ and $0<q<1$ are set numbers.
+where $Q>1$ and $0<q<1$ are parameters.
+%Russian
+Значение параметра $Q$ определяет наши априорные предположения о том, во сколько раз относительная разность $\mu_i$, вычисленная на интервале $(x_{i-1},x_i)$, содержащем точку разрыва, будет превышать максимальную разность $\mu$, найденную по всем интервалам, принадлежащим области непрерывности функции.
+Данное значение задается исследователем, и в проведенных эксмериментах оно было выбрано равным 3.
 
@@
 \end{enumerate}
+
+\textbf{Remark}.
+В соответствии с (\ref{gamma}) and (\ref{delta_cond}) значение $\gamma_i+\gamma_{i-1}+|\delta_i |$ может быть равно 0 либо 1, в зависимости от наличия либо отсутствия точки разрыва в интервале $(x_{i-1},x_i)$.
+Так, если данный интервал не содержит точек разрыва, т.е. $\gamma_i+\gamma_{i-1}+|\delta_i | = 0$, его характеристика (\ref{R_i}) пример вид 
+\[
+R(i)=\Delta_i+\frac{ (z_i-z_{i-1} )^2}{(r\mu)^2 \Delta_i }-\frac{2(z_i + z_{i-1})}{r\mu}.
+\]
+Если одна из граничных точек интервала является заданной точкой разрыва, т.е. $\gamma_{i-1} = 1$ или $\gamma_i = 1$, то мы получим соответственно характеристики
+\begin{equation}\label{RR}
+R(i)=2\Delta_i - \frac{4 z_i }{r\mu}, \; R(i)=2\Delta_i - \frac{4 z_{i-1} }{r\mu}.
+\end{equation}
+Аналогичные формулы для характеристик интервалов будут получаться, если $(x_{i-1},x_i)$ является интервалом, который содержит незаданную точку разрыва, т.е.
+когда $|\delta_i | = 1 $.
+В этом случае мы таже будем получать формулы (\ref{RR}) для $\delta_i = -1$ и $\delta_i = 1$ соответственно.
 

### 2021-09-24 2021/Entropy/entropy.tex
Translate aswers for reviewers

--- 2021-09-19 2021/Entropy/entropy.tex
@@
 %Russian
-Методы машинного обучения активно используются в сочетании с алгоритмами bayesian optimization, основанными на использовании probabilistic surrogate models of the objective function.
-Детальный обзор данного направления развития методов глобальной оптмизации представлен в \cite{Archetti2019,Zhigljavsky2021}.
+{ \color{red}
+Machine learning methods are actively used in combination with bayesian optimization algorithms based on the probabilistic surrogate models of the objective function.
+The detailed overview of this trend in the development of global optimization methods is presented in \cite{Archetti2019,Zhigljavsky2021}.
+}
 
@@
 
-Note that the algorithms of the last class do not provide guaranteed convergence to the solution of the problem and are inferior to deterministic algorithms in terms of the quality of solution \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a particular set).
+Note that the algorithms of the latter class do not provide guaranteed convergence to the solution of the problem and are inferior to deterministic algorithms in terms of the quality of solution \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a particular set).
 Therefore, deterministic methods seem to be potentially more effective.
@@
 %Russian
-Настоящая работа направлена на дальнейшее развитие эффективного детерминированного метода глобальной оптимизации, известного как \textit{the information-statistical global search algorithm} \cite{Strongin2000}.
-Указанная здесь книга содержит результаты теоретических исследований метода, имеющих непосредственное значение для его практической реализации.
-В частности, в ней детально рассмотрены вопросы сходимости, выбора параметров и условия остановки алгоритма, и т.д.
-Отметим, что алгоритм глобального поиска первоначально был ориентирован на решение unconstrained optimization problems.
-Впоследствии он был обобщен для решения problems with non-convex constraints\cite{Barkalov2002} and multicriteria optimization problems \cite{Gergel2020}.
-Одновременно с этим были предложены различные параллельные варианты указанных алгоритмов, которые могут быть использованы на современных суперкомпьютерах \cite{Barkalov2016,globalizerSystem,Strongin2018}.
+This paper aims to further develop the efficient deterministic global optimization method known as \textit{the information-statistical global search algorithm} \cite{Strongin2000}.
+The book referenced here contains the results of theoretical studies of the method that are of direct importance for its practical implementation.
+In particular, it discusses in detail the issues of convergence, the choice of parameters, and the conditions for stopping the algorithm, etc.
+Note that the global search algorithm was originally designed for solving unconstrained optimization problems.
+Later, it was generalized to solve problems with non-convex constraints\cite{Barkalov2002} and multicriteria optimization problems \cite{Gergel2020}.
+At the same time, scholars proposed various parallel versions of these algorithms, which can be used on modern supercomputers.\cite{Barkalov2016,globalizerSystem,Strongin2018}.
 
@@
 %Russian
-В целом, можно сказать, что дерево принятия решений -- это бинарное дерево, в листьях которого находятся значения функции, а в остальных узлах -- условия перехода.
-В нашем случае при применении дерева решений для построения регрессии каждому узлу соответствует несколько результатов испытаний и значение узла $c_j$, которое вычисляется как среднее квадратичное значений в точках испытаний, приписанных к этому узлу.
-Итоговая кусочно-постоянная аппроксимация строится с использованием значений $c_j$, которые находятся в листьях дерева.
+{\color{red}
+Generally speaking, a decision tree is a binary tree, the leave nodes of which contain the values of the function, and the other nodes contain the transition conditions.
+In our case, when applying a decision tree to construct a regression, each node corresponds to the results of several trials and the value $c_j$, which is calculated as the mean square of the trial points assigned to this node.
+The final piecewise constant approximation is constructed using the values $c_j$, located in the leave nodes of the tree.
 
-Дерево строится рекурсивно, начиная с корневого узла. 
-В каждом узле применяется правило принятия решения: все данные  разбиваются на две группы (согласно некоторому правилу разбиения) и отправляются в левый и правый узлы-потомки.
-Затем процедура рекурсивно разделяет левый и правый узлы.
-Остановка рекурсии в узле происходит в одном из следующих случаев:
+The tree is built recursively, starting from the root node. 
+A decision rule is applied at each node: all data is divided into two groups (according to a partitioning rule) and sent to the left and right child nodes.
+The procedure then recursively separates the left and right nodes.
+Recursion stops at a node in one of the following cases:
 \begin{itemize}
-	\item Количество приписанных к узлу точе испытаний становится меньше указанного порогового значения (мы использовали 1).
-	\item Сумма квадратов отклонений значений функции от значения $c_j$, приписанного к этому узлу, становится меньше заданной точности (мы использовали 0.0001).
+	\item The number of trial points assigned to the node becomes less than the specified threshold value (we used 1).
+	\item The sum of the squared deviations of the function values from the value $c_j$, assigned to this node becomes less than the set accuracy (we used $10^{-3}$).
 \end{itemize}
-Подробнее об алгоритме построения регрессии с использованием деревьев решений см., например, \cite{Breiman1984}.
+For more information on the algorithm for constructing regression using decision trees, see, for example \cite{Breiman1984}.
+}
 
-Регрессия, построенная с помощью деревьев решений,
+Regression built using decision trees,
 is, on the one hand, rather simple, and on the other hand, it adequately reflects the properties of the function under investigation (the presence or absence of local minima).
@@
 %Russian
-Использование additional attribute $q^i, \; 0 \leq i \leq k ,$ позволяет различать точки, полученные во время глобального ($q^i=0$) и во время локального ($q^i=1,2$) поиска.
-Интервалы, граничные точки которых имеют значение $q^i=1$ или $q^i=2$, будут пропускаться при глобальном поиске, т.е.
-в них не будут проводиться дальнейшие испытания.
-А выделенное значение $q^i=2$, соответствующее локальному минимуму, будет позволять использовать данные точки при проверке критерия остановки.
-Если по правилам глобального поиска новая точка испытания должна быть поставлена в заданной $\epsilon$-окрестности одного из найденных локальных минимумов, это будет соответствовать окончанию поиска.
+{\color{red}
+Using the additional attribute $q^i, \; 0 \leq i \leq k ,$ makes it possible to distinguish between points obtained during a global ($q^i=0$) and a local ($q^i=1,2$) search.
+The intervals, the boundary points of which have the value  $q^i=1$ or $q^i=2$, will be skipped in the global search, i.e.
+they will not be tried further.
+While the selected value $q^i=2$, which corresponds to the local minimum will allow using these points for checking the stopping criterion.
+When the rules of the global search signal that a new trial point needs be placed in a given  $\epsilon$-neighborhood of one of the found local minima, this will correspond to the end of the search.
+}
 
@@
 %Russian
-Отметим, что указанный критерий остановки проверяется после Step 6, т.е.
-после вычисления по правилам глобального поиска точки $x^{k+1}$ очередного испытания, но до проведения самого испытания в ней.
-Смысл данного критерия заключается в следующем.
-Поиск останавиливается, когда интервал $(x_{t-1},x_t)$, в который попадает точка $x^{k+1}$, становится достаточно малым, либо когда точка $x^{k+1}$ попадает в малую окрестность одного из найденных локальных минимумов, который в этом случае будет одновременно являться и глобальным.
+{\color{red}
+Note that the specified stopping criterion is checked after Step 6, i.e.
+after the calculation according to the rules of the global search for the point of the next trial $x^{k+1}$, but before the trial itself is carried out in it.
+The idea behind this criterion is as follows.
+The search stops when the interval  $(x_{t-1},x_t)$, in which the point $x^{k+1}$ falls, becomes sufficiently small, or when the point $x^{k+1}$ falls into a small neighborhood of one of the found local minima.
+Such minima in this case will be simultaneously global.
+}
 
@@
 Let us compare the basic global search algorithm (GSA) and its decision tree-based modification (GSA-DT) with the well-known DIRECT global optimization algorithm  \cite{Jones2009}.
+{\color{red}
 %Russian 
-Выбор именного этого метода для сравнения объясняется следующим.
-DIRECT является одним из наиболее известных и популярных детерминированных методов решения задач глобальной оптмизации with ``black box'' objective function.
-Обзор различных модификаций метода, а также примеров решения задач приведен в \cite{Jones2021}.
-Для этого метода известно, что при достаточно большом числе поисковых испытаний DIRECT гарантированно найдет глобальное решение задачи.  
-Однако если использовать критерий основки по точности, то метод может быстро остановиться в одном из локальных минимумов.
-Сказанное подтверждается приведенными в данном разделе результатами экспериментов.
+The choice of this particular method for comparison is explained as follows.
+DIRECT is one of the most well-known and popular deterministic methods for solving global optimization problems with a `` black box '' objective function.
+An overview of various modifications of the method, as well as examples of solving problems, is given in \cite{Jones2021}.
+It is known that with a sufficiently large number of search trials, DIRECT is guaranteed to find a global solution to the problem. 
+However, if the stopping by accuracy is used as a criterion, then the method can abruptly stop at one of the local minima, which is confirmed by the experimental results presented in this section.
 
-Если же говорить про детерминированные оптимизационные алгоритмы, реализованные in popular Computer Algebra Systems (CAS), то их сравнение с такими методами, как DIRECT или GSA, будет некорректным по следующей причине.
-Оптимизационные алгоритмы из CAS ориентированы на решение задач с целевой функцией, заданной явно, в виде формулы.
-Формульное задание целевой функции предполагает известным также и ее производные, что позволяет использовать методы первого порядка, обладающие значительно более быстрой сходимостью, чем методы нулевого порядка, к которым отноится и DIRECT, и GSA.
-Но при этом методы первого порядка не гарантируют нахождение глобального минимума.
-Например, система Mathematica предлагает для решения задач глобальной оптмизации методы, которые гарантированно найдут глобальное решение задачи только в случае, если целевая функция и ограничения являются линейными или выпуклыми.
+Regarding the deterministic optimization algorithms implemented in popular Computer Algebra Systems (CAS), their comparison with methods such as DIRECT or GSA will be incorrect for the following reason.
+Optimization algorithms from CAS are focused on solving problems with an objective function set explicitly, in the form of a formula.
+Formula definition of the objective function assumes that its derivatives are also known, which makes it possible to use first-order methods with significantly faster convergence than zero-order methods, which include both DIRECT and GSA.
+However, first-order methods do not guarantee finding the global minimum.
+For example, the Mathematica system offers methods for solving global optimization problems that are guaranteed to find a global solution to the problem only if the objective function and constraints are linear or convex.
 Otherwise, the result may sometimes only be a local minimum.
 
-Более того, методы, требующие формульного задания целевой функции, не могут применяться к решению большого класса прикладных задач, в которых вид целевой функции не известен, а ее значения вычисляются в результате численного моделирования.
-Использование для решения задач подобного вида эвристических методов (таких как Differential Evolution, Simulated Annealing or Random Search) также не всегда приводит к решению задачи.
-These methods often do find the global minimum, but are not guaranteed to do so. В терминах числа корректно решенных задач эвристические методы проигрывают детерминированным \cite{Sergeyev2018}.
+Moreover, methods that require a formulaic specification of the objective function cannot be applied to the solution of a large class of applied problems in which the form of the objective function is not known, and its values are calculated as a result of numerical simulation.
+The use of heuristic methods for solving problems of this kind (such as Differential Evolution, Simulated Annealing or Random Search) also does not always lead to the solution of the problem.
+These methods often do find the global minimum, but are not guaranteed to do so.
+In terms of the number of correctly solved problems, heuristic methods are inferior to deterministic ones \cite{Sergeyev2018}.
+}
 
@@
 The number of unsolved problems is indicated in parentheses.
-%Russian
-В этих и последующих таблицах для метода GSA-DT приведено полное число испытаний, которые были выполнены как во время глобального, так и во время локального поиска.
-При решении задач с точностью $10^{-2}$, $10^{-3}$, $10^{-4}$ число запусков локального поиска при решении одной задачи составляло в среднем 2, 3 и 4 соответственно; а отношение числа испытаний, выполненных по правилам глобального поиска, к числу испытаний, выполненных локальным методом, составляло (при соответствующей точности) 7.2, 6.0, 4.5 для задач класса $F_{HL}$ и 3.2, 2.3, 1.9 для задач класса $F_{SH}$.
+{\color{red}
+These and subsequent tables feature the total number of trials that were performed during both global and local searches by GSA-DT.
+When solving problems with an accuracy  $10^{-2}$, $10^{-3}$, $10^{-4}$ the number of local search launches when solving one problem on average was equal to 2, 3, and 4, respectively; and the ratio of the number of trials performed according to the global search rules to the number of trials performed by the local method was (with appropriate accuracy) 7.2, 6.0, 4.5 for problems of the $F_{HL}$ class and 3.2, 2.3, 1.9 for the $F_{SH}$ class problems.
+}
 
@@
 %Russian
-Еще одним направлением дальнейшей работы будет являться комбинирование предложенного подхода к использованию локального поиска с традиционными способами ускорения алгоритмов глобальной оптмизации.
-Например, широко известным способом ускорения алгоритмов данного класса является использование local adaptive estimates of the Lipschitz constant $L_i$ в различных зонах $D_i \in D$ области поиска $D$ вместо одной глобальной оценки константы $L$ во всей области $D$ ( see \cite{Kvasov2020,Sergeyev2021}).
-Это позволяет алгоритму настраиваться на локальное поведение целевой функции в различных частях допустимой области и сокращает число поисковых испытаний, требующихся для достижения сходимости.
-Использование только глобальной информации о поведении целевой функции в ходе ее минимизации может значительно замедлить сходимость алгоритма к точке
-глобального минимума.
-В данном направлении также планируется использовать методы машинного обучения для выделения подобластей $D_i \in D$ и построения локальных оценок константы Липшица $L_i$ в выделенных подобластях.
+{\color{red}
+Another direction for further work will be the combination of the proposed approach to the application of local search with traditional methods of accelerating global optimization algorithms.
+For example, a well-known way to speed up algorithms of this class is to use local adaptive estimates of the Lipschitz constant $L_i$ in various subdomains $D_i \in D$ within the search domain $D$ instead of a single global estimate of the constant $L$ for the entire domain $D$ ( see \cite{Kvasov2020,Sergeyev2021}).
+This allows the algorithm to adjust to the local behavior of the objective function in different parts of the feasible domain, and, thereby, reduces the number of search trials required to achieve convergence.
+Using only global information about the behavior of the objective function during its minimization can significantly slow down the convergence of the algorithm to a point of global minimum.
+Further research on this issue could focus on using machine learning methods to isolate subdomains $D_i \in D$ and construct local estimates of the Lipschitz constant  $L_i$ in the selected subdomains.
+}
 

### 2021-09-25 2021/Entropy/entropy.tex
First revision

@@
 Later, it was generalized to solve problems with non-convex constraints\cite{Barkalov2002} and multicriteria optimization problems \cite{Gergel2020}.
-At the same time, scholars proposed various parallel versions of these algorithms, which can be used on modern supercomputers.\cite{Barkalov2016,globalizerSystem,Strongin2018}.
+At the same time, scholars proposed various parallel versions of these algorithms, which can be used on modern supercomputers \cite{Barkalov2016,globalizerSystem,Strongin2018}.
 

### 2021-09-29 2021/ICCS_BarkalovUsova/iccs.tex
Some comments in Russian

@@
 %Russian
-
 \textbf{Remark 1}.
@@
 To eliminate the effect of small relative differences of intervals containing extreme points, the additional constraint $1 \leq p < q(m-2(s+1))$ is introduced in (\ref{mu_main_cond}) to exclude small values of $\mu_i$ from consideration.
+%Russian
+Значение параметра $q$ отражает априорные предполжения о том, какая доля поисковых интервалов в их общем числе не будет содержать точек локальных экстремумов.
 
@@
 
+%Russian
 \textbf{Remark}.
@@
 когда $|\delta_i | = 1 $.
-В этом случае мы таже будем получать формулы (\ref{RR}) для $\delta_i = -1$ и $\delta_i = 1$ соответственно.
+В этом случае мы таже будем получать формулы (\ref{RR}) для $\delta_i = -1$ и $\delta_i = 1$, что соответствует использованю результатов испытания в правой точке интервала для построения его характеристики в случае убывания функции и выбору левой точки при возрастании функции.
 

### 2021-09-30 2021/ICCS_BarkalovUsova/iccs.tex
some corrections in Russian

--- 2021-09-24 2021/ICCS_BarkalovUsova/iccs.tex
@@
 \textbf{Remark 2}.
-Приведенные выше формулы для вычисления характеристик интервалов могут быть использованы (в модифицированной форме) в алгоритме решения задач невыпуклыми ограничениями \cite{Sergeyev1995,Sergeyev2006_1,Sergeyev2007}.
+Приведенные выше формулы для вычисления характеристик интервалов могут быть использованы (в модифицированной форме) в алгоритме решения задач невыпуклыми ограничениями \cite{Barkalov2002,Sergeyev2006_1,Sergeyev2007,Sergeyev1995}.
 Здесь решение задачи с частично определенными ограничениями неявно сводится к минимизации разрывной функции, в которой точками разрыва являются границы допустимого множества решений задачи.
@@
 %Russian
-Значение параметра $Q$ определяет наши априорные предположения о том, во сколько раз относительная разность $\mu_i$, вычисленная на интервале $(x_{i-1},x_i)$, содержащем точку разрыва, будет превышать максимальную разность $\mu$, найденную по всем интервалам, принадлежащим области непрерывности функции.
+Значение параметра $Q$ определяет априорные предположения о том, во сколько раз относительная разность $\mu_i$, вычисленная на интервале $(x_{i-1},x_i)$, содержащем точку разрыва, будет превышать максимальную разность $\mu$, найденную по всем интервалам, принадлежащим области непрерывности функции.
 Данное значение задается исследователем, и в проведенных эксмериментах оно было выбрано равным 3.
@@
 
+%Russian
 \textbf{Remark}.
 В соответствии с (\ref{gamma}) and (\ref{delta_cond}) значение $\gamma_i+\gamma_{i-1}+|\delta_i |$ может быть равно 0 либо 1, в зависимости от наличия либо отсутствия точки разрыва в интервале $(x_{i-1},x_i)$.
-Так, если данный интервал не содержит точек разрыва, т.е. $\gamma_i+\gamma_{i-1}+|\delta_i | = 0$, его характеристика (\ref{R_i}) пример вид 
+Так, если данный интервал не содержит точек разрыва, т.е. $\gamma_i+\gamma_{i-1}+|\delta_i | = 0$, то его характеристика (\ref{R_i}) примет вид 
 \[
@@
 когда $|\delta_i | = 1 $.
-В этом случае мы таже будем получать формулы (\ref{RR}) для $\delta_i = -1$ и $\delta_i = 1$ соответственно.
+В этом случае мы таже будем получать формулы (\ref{RR}) для $\delta_i = -1$ и $\delta_i = 1$, что соответствует использованию результатов испытания в правой точке интервала для построения его характеристики в случае убывания функции и выбору левой точки при возрастании функции.
 

### 2021-10-01 2021/ICCS_BarkalovUsova/iccs.tex
one more correction in Russian

@@
 This algorithm was developed on the basis of the \textit{global search algorithm} proposed by Strongin \cite{Strongin2000} and provides a search for the global optimum.
+%Russian
+Проводимое в статье рассмотрение ограничено одномерным случаем.
+Обобщение нового алгоритма на многомерный случай может быть выполнено в соответствии со ставшими уже стандартными подходами из \cite{Grishagin2016,Grishagin2018}, использующими adaptive nested optimization scheme.
+Указанная схема редукции размерности позволяет заменить решение исходной многомерной задачи решением семейства одномерных задач, рекурсивно связанных между собой.
+
+Основная часть статьи организована следующим образом.
+
 Section 2 describes the algorithm for functions with given discontinuities.
@@
 %Russian
-Значение параметра $q$ отражает априорные предполжения о том, какая доля поисковых интервалов в их общем числе не будет содержать точек локальных экстремумов.
+Параметр $0<q<1$ тажке задается исследователем.
+Значение $1 - q$ отражает априорные предполжения о том, какая доля поисковых интервалов будет иметь значения $\mu_i$ близкими к нулю, т.е.
+фактически характеризует степень многоэкстремальности задачи.
+В проведенных экспериментах использовалось значение $q=0.9$.
 

### 2021-10-02 2021/ICCS_BarkalovUsova/iccs.tex
final answers in Russian

@@
 Проводимое в статье рассмотрение ограничено одномерным случаем.
-Обобщение нового алгоритма на многомерный случай может быть выполнено в соответствии со ставшими уже стандартными подходами из \cite{Grishagin2016,Grishagin2018}, использующими adaptive nested optimization scheme.
+Обобщение нового алгоритма на многомерный случай может быть выполнено в соответствии со ставшими уже стандартными подходами из \cite{Grishagin2016,Grishagin2018}, использующими редукцию размерности на основе adaptive nested optimization scheme.
 Указанная схема редукции размерности позволяет заменить решение исходной многомерной задачи решением семейства одномерных задач, рекурсивно связанных между собой.
@@
 Еще раз отметим, что в соответствии с (\ref{gamma}) значение $\gamma_i+\gamma_{i-1}$ может быть равно либо 0 (обе граничных точки интервала принадлежат области непрерывности функции), либо 1 (одна из граничных точек является точкой разрыва).
-В этом случае формула (\ref{R_i}) иметь вид
+В первом случае формула (\ref{R_i}) будет иметь вид
 \[
-R(i)=\Delta_i+\frac{ (z_i-z_{i-1} )^2}{(r\mu)^2 \Delta_i }-\frac{2(z_i + z_{i-1})}{r\mu},
+R(i)=\Delta_i+\frac{ (z_i-z_{i-1} )^2}{(r\mu)^2 \Delta_i }-\frac{2(z_i + z_{i-1})}{r\mu}.
 \]
-в случае отстутсвия точке разрыва и 
+А во втором случае будем иметь
 \[
@@
 \textbf{Remark 2}.
-Приведенные выше формулы для вычисления характеристик интервалов могут быть использованы (в модифицированной форме) в алгоритме решения задач невыпуклыми ограничениями \cite{Barkalov2002,Sergeyev2006_1,Sergeyev2007,Sergeyev1995}.
-Здесь решение задачи с частично определенными ограничениями неявно сводится к минимизации разрывной функции, в которой точками разрыва являются границы допустимого множества решений задачи.
+Приведенные выше формулы для вычисления характеристик интервалов могут быть использованы (в модифицированной форме) в алгоритме решения задач with non-convex constraints \cite{Barkalov2002,Sergeyev2006_1,Sergeyev2007,Sergeyev1995}.
+Здесь решение задачи with partially defined constraints неявно сводится к минимизации разрывной функции, в которой точками разрыва являются границы допустимого множества решений задачи.
 
@@
 фактически характеризует степень многоэкстремальности задачи.
-В проведенных экспериментах использовалось значение $q=0.9$.
+В проведенных нами экспериментах использовалось значение $q=0.9$.
 

### 2021-10-10 2021/ICCS_BarkalovUsova/optima.tex
After answers for the reviewers

--- 2021-09-30 2021/ICCS_BarkalovUsova/iccs.tex
@@
 %%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
-\usepackage{inputenc}
-\usepackage[english,russian]{babel}
-\usepackage{cmap}
+%\usepackage{inputenc}
+%\usepackage[english,russian]{babel}
+%\usepackage{cmap}
 %%%%
@@
 %
-\title{A search algorithm for the global extremum \break of a discontinuous function\thanks{This work was supported by the Russian Science Foundation, project No.\,21-11-00204.}}
+\title{A Search Algorithm for the Global Extremum \break of a Discontinuous Function\thanks{This work was supported by the Russian Science Foundation, project No.\,21-11-00204.}}
 %
-\titlerunning{A search algorithm for the global extremum}
+\titlerunning{A Search Algorithm for the Global Extremum}
 % If the paper title is too long for the running head, you can set 
@@
 %
-\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, 
+\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, 
 Russia \email{konstantin.barkalov@itmm.unn.ru}}
@@
 \begin{abstract}
-We consider the problem of finding the global minimum of a one-dimensional function that may have several finite jump discontinuity points.
-The discontinuities of the objective function can reflect the specifics of the optimized object in the mathematical model (for example, shock effects, resonance phenomena, jumps in geometric dimensions or material properties, etc.).
-In some cases, the discontinuing set is known.
-However, at the same time, there are problems in which there are no a priori estimates of discontinuity points, but it is known that such points are possible.
-The paper gives a description of the global search algorithm for solving the specified class of problems.
-The authors conducted an experimental comparison of the proposed algorithm with the methods implemented in MATLAB Global Optimization Toolbox.
+This article discusses the problem of finding the global minimum of a one-dimensional function that may have several finite jump discontinuity points.
+The discontinuities of the objective function may be due to the nature of the optimized object in the mathematical model (for example, shock effects, resonance phenomena, jumps in geometric dimensions or material properties, etc.).
+In some cases, the discontinuity poits are known; but at the same time, there are problems with no a priori estimates of discontinuity points.
+It is known, however, that such points do exist.
+The paper gives a description of the global search algorithm for solving such class of problems.
+In addition, the authors conducted an experimental comparison of the proposed algorithm with the methods of MATLAB Global Optimization Toolbox.
 
@@
 
-Global optimization problems involving many local extrema arise in many areas of applications (see, e.g., \cite{Cavoretto2021,Kvasov2013,Kvasov2015,Modorskii2016}).
-As a rule, the calculation of even a single value of the objective function in such problems requires considerable time, since the analytical form of the function is not known and the calculation of its values involves time-consuming numerical modelling.
-In this case, we have to deal with expensive black-box optimization problems as they are often called.
-Problems of this class are extremely complicated and are the object of intensive research.
+Global optimization problems involving many local extrema arise in many applications (see, e.g.
+\cite{Cavoretto2021,Kvasov2013,Kvasov2015,Modorskii2016}).
+As a rule, the calculation of even a single value of the objective function in these problems takes considerable time, since the analytical form of the function is not known and the calculation of its values involves time-consuming numerical modelling.
+That is why these computationally expensive optimization problems are often called black-box problems.
+They are extremely complicated and are the subject of intensive research.
 
-One should note an important role of one-dimensional global optimization problems.
-Firstly, as has already been mentioned, they arise in many applications (see, e.g., \cite{Gillard2017,Sergeyev2020}).
-Secondly, many approaches to solving multidimensional optimization problems are, in one way or another, based on reducing the solution of the original multidimensional problem to solving a series of related one-dimensional optimization problems (see, for example, popular approaches in \cite{PaulaviciusZilinskas2014,Pinter1996,Sergeyev2017,Sergeyev2013}).
+Despite their complex nature, these problems play an important role in many aspects of mathematical research.
+First, as has already been mentioned, they are applied for multiple tasks in a variety of fields (see, e.g., \cite{Gillard2017,Sergeyev2020}).
+Secondly, many approaches to solving multidimensional optimization problems, in one way or another, are based on reducing the solution of the original multidimensional problem to solving a series of related one-dimensional optimization problems (see, for example, popular approaches in \cite{PaulaviciusZilinskas2014,Pinter1996,Sergeyev2017,Sergeyev2013}).
 
 The literature describes various methods for solving the problem (\ref{problem}) depending on the assumptions about the properties of the objective function \cite{Horst1995,Horst1996,Pinter1996}.
-One of such assumptions, which is often fulfilled in applied optimization problems, is that the objective function $\varphi(x)$ satisfies the Lipschitz condition
+One of such assumptions, which is often made about  applied optimization problems, is that the objective function $\varphi(x)$ satisfies the Lipschitz condition
 \[
@@
 where the constant $L$ is unknown a priori. 
-This feature of the objective function is commonly used in many approaches to the development of algorithms for multiextremal optimization \cite{Evtushenko2009,Evtushenko2013,Jones2009,Paulavicius2011}.
+This feature of the objective function is used for the development of algorithms for multiextremal optimization \cite{Evtushenko2009,Evtushenko2013,Jones2009,Paulavicius2011}.
 
-However, in some applied problems, the Lipschitz property may not be fulfilled due to the presence of discontinuous jumps in the values of the function at some points of the search domain.
-Such jumps can be caused by sharp changes in the characteristics of the optimized object with small changes in its parameters, for example:
+However, in some applied problems, the Lipschitz condition may not be satisfied due to the presence of discontinuous jumps in the values of the function at certain points of the search domain.
+Such jumps can be caused by sharp changes in the characteristics of the optimized object which trigger minor changes in its parameters, for example:
 \begin{itemize}
-  \item changes in the thickness of the material (or other geometric parameters of the object);
-  \item changes in material properties (both when different materials are used in one design, and with changes in the properties of the same material as a result of technological operations, for example, welding, etc.);
+  \item changes in the thickness of material (or other geometric parameters of the object);
+  \item changes in physical properties (both when different materials are used in a single design project, and when the properties of one material change as a result of technological operations, for example,  welding, etc.);
   \item changes in external sources of impact (temperature, force, etc.).
@@
 For these cases, the set of points, at which the characteristics of the object jump, is typically known in advance.
-However, at the same time, there are problems in which there are no a priori estimates of discontinuity points, but it is known that such points are possible.
+However, there are problems with no a priori estimates of discontinuity points.
+It is known, however, that such points may exist.
 
 The known methods for solving such problems, as a rule, either generalize the concept of a gradient for discontinuous functions \cite{Batukhtin1993,Batukhtin1998,Moreau2000}, or belong to the class of bio-inspired algorithms \cite{Ban2019,ZhangXu}.
-These methods provide, generally speaking, the search for only a local solution to the problem.
+Generally speaking, these methods only search for a local solution to the problem.
 
 In this paper, we will consider a deterministic algorithm for solving global optimization problems with both given and unspecified discontinuity points.
-This algorithm was developed on the basis of the \textit{global search algorithm} proposed by Strongin \cite{Strongin2000} and provides a search for the global optimum.
+This algorithm was developed on the basis of the \textit{global search algorithm} proposed by Strongin \cite{Strongin2000}, and it can be used to find the global optimum.
+%Russian
+The scope of this article is limited to the one-dimensional case.
+Generalization of the new algorithm to the multidimensional case can be performed in accordance with the already existing standard approaches from \cite{Grishagin2016,Grishagin2018}, which use a dimensionality reduction based on the adaptive nested optimization scheme.
+The indicated dimensionality reduction scheme allows replacing the solution of the original multidimensional problem with the solution of a family of one-dimensional problems that are recursively related to each other.
+
+The main part of the article is organized as follows.
 Section 2 describes the algorithm for functions with given discontinuities.
 Section 3 discusses how to identify unspecified discontinuities.
-Section 4 describes how this discontinuity detection method can be used in the original algorithm.
+Section 4 describes how the discontinuity detection method can be used in the original algorithm.
 Section 5 presents the results of a comparison of the developed algorithm with the global search methods implemented in MATLAB Global Optimization Toolbox \cite{MatlabOTB}.
@@
 
-\section{Algorithm for functions with given discontinuity points}
+\section{Algorithm for Functions with Given Discontinuity Points}
 
-Let us consider problem (\ref{problem}) assuming that the objective function $\varphi(x), x\in[a,b],$ has jump discontinuities at the given points
+Let us consider the problem (\ref{problem}) assuming that the objective function $\varphi(x), x\in[a,b],$ has jump discontinuities at the given points
 \begin{equation}\label{points}
@@
 \end{equation}
-and satisfies the Lipschitz condition between them, i.e.
+and satisfies the Lipschitz condition between these points, i.e.
 \begin{equation}\label{LipschitzСondition}
@@
 \end{equation}
-Points $\omega_{0}$ and $\omega_{s+1}$ are included in the general list regardless of the presence of function discontinuities in them for the convenience of subsequent presentation.
+Points $\omega_{0}$ and $\omega_{s+1}$ are included in the general list regardless of the presence of function discontinuities at them for the convenience of subsequent presentation.
 Let us denote the resulting set of points $\Omega$, i.e.
@@
 
-In \cite{Strongin2000}, the global search algorithm (GSA) was considered for solving problems without discontinuities.
+In \cite{Strongin2000}, the global search algorithm (GSA) was applied for solving problems without discontinuities.
 The authors also proposed a method for accounting for discontinuities based on a smoothing transformation of the objective function.
-Below we consider an algorithm based on GSA, in which the discontinuous function is minimized without smoothing of the discontinuities, which greatly simplifies the computational scheme of the algorithm.
+Below we consider an algorithm based on GSA, in which the discontinuous function is minimized without smoothing of the discontinuities, which substantially simplifies the computational scheme of the algorithm.
 
@@
 
-\item Check termination condition $x_t-x_{t-1} \leq \epsilon$, where $t$ is the number of the interval with the maximal characteristic and $\epsilon > 0$ is the predefined accuracy.
+\item Check the stopping condition $x_t-x_{t-1} \leq \epsilon$, where $t$ is the number of the interval with the maximal characteristic and $\epsilon > 0$ is the predefined accuracy.
 
@@
 \textbf{Remark 1}.
-Еще раз отметим, что в соответствии с (\ref{gamma}) значение $\gamma_i+\gamma_{i-1}$ может быть равно либо 0 (обе граничных точки интервала принадлежат области непрерывности функции), либо 1 (одна из граничных точек является точкой разрыва).
-В этом случае формула (\ref{R_i}) иметь вид
+In accordance with (\ref{gamma}), the value $\gamma_i+\gamma_{i-1}$ can be equal to either 0 (both boundary points of the interval belong to the domain of continuity of the function), or 1 (one of the boundary points is the jump point).
+In the first case, the formula  (\ref{R_i}) will have the form
 \[
-R(i)=\Delta_i+\frac{ (z_i-z_{i-1} )^2}{(r\mu)^2 \Delta_i }-\frac{2(z_i + z_{i-1})}{r\mu},
+R(i)=\Delta_i+\frac{ (z_i-z_{i-1} )^2}{(r\mu)^2 \Delta_i }-\frac{2(z_i + z_{i-1})}{r\mu}.
 \]
-в случае отстутсвия точке разрыва и 
+And in the second case we will have
 \[
@@
 \]
-в случае разрыва в точке $x_{i-1}$ или $x_{i}$ соответственно.
+in case of discontinuity at the point  $x_{i-1}$ or $x_{i}$ respectively.
 
 \textbf{Remark 2}.
-Приведенные выше формулы для вычисления характеристик интервалов могут быть использованы (в модифицированной форме) в алгоритме решения задач невыпуклыми ограничениями \cite{Barkalov2002,Sergeyev2006_1,Sergeyev2007,Sergeyev1995}.
-Здесь решение задачи с частично определенными ограничениями неявно сводится к минимизации разрывной функции, в которой точками разрыва являются границы допустимого множества решений задачи.
+The above formulas for calculating the characteristics of intervals can be used (in a modified form) in the algorithm for solving problems with non-convex constraints  \cite{Barkalov2002,Sergeyev2006_1,Sergeyev2007,Sergeyev1995}.
+Here, the solution to the problem with partially defined constraints is implicitly reduced to minimizing a discontinuous function, in which the discontinuity points are the boundaries of the admissible set of the constrained problem.
 
@@
 
-It can be proved that the conditions for the convergence of this algorithm will correspond to the conditions for the convergence of its basic prototype -- the global search algorithm from \cite{Strongin2000}.
+It can be proved that the convergence conditions for this algorithm correspond to the convergence conditions for its prototype -- the global search algorithm from \cite{Strongin2000}.
 The formulation and rigorous proof of the convergence conditions will be the subject of further publications.
 
-Let us illustrate the operation of the algorithm when searching for the minimum of the function
+Let us illustrate the operation of the algorithm during the search for the minimum of the function
 \begin{equation}\label{testFunction}
@@
 at $x\in[0,10]$.
-All discontinuity points were considered known, the parameters of the method $r=2.0,\epsilon=0.001$ were used.
-Fig.~\ref{ris1} shows a graph of this function, the discontinuities are marked with a dotted line.
+Here, all the discontinuity points were considered to be known, and the parameters of the method were $r=2.0,\; \epsilon=0.001$.
+Fig.~\ref{ris1} shows a graph of this function; the discontinuities are marked with a dotted line.
 The dashes under the graph indicate the points of 70 trials required for the algorithm to solve problem with the specified accuracy.
@@
 
-Let us suppose now that the location of the discontinuity points for the function $\varphi(x)$ is not known, but it is known that such points are possible.
-Let the function have a discontinuity at an a priori unknown point $\omega \in(x_{i-1},x_i)$, then the relative difference $\mu_i$ from (\ref{mu_i}) will significantly exceed the relative differences for other intervals since within them the function satisfies the Lipschitz condition.
-Therefore, the identification of such situations can be based on a comparison of the relative differences $\mu_i$ from (\ref{mu_i}) corresponding to different search intervals $(x_{i-1},x_i)$.
+Let us suppose that the location of the discontinuity points for the function $\varphi(x)$ is not known, but it is known that such points are possible.
+Let the function have a discontinuity at an a priori unknown point $\omega \in(x_{i-1},x_i)$, then the relative difference $\mu_i$ from (\ref{mu_i}) will significantly exceed the relative differences for other intervals, since within within these intervals the function satisfies the Lipschitz condition.
+Therefore, such instances can be identified based on a comparison of the relative differences $\mu_i$ from (\ref{mu_i}) corresponding to different search intervals $(x_{i-1},x_i)$.
 
@@
 %Russian
-Значение параметра $Q$ определяет априорные предположения о том, во сколько раз относительная разность $\mu_i$, вычисленная на интервале $(x_{i-1},x_i)$, содержащем точку разрыва, будет превышать максимальную разность $\mu$, найденную по всем интервалам, принадлежащим области непрерывности функции.
-Данное значение задается исследователем, и в проведенных эксмериментах оно было выбрано равным 3.
+The value of the parameter $Q$ determines the a priori assumptions about how many times the relative difference $\mu_i$, calculated on the interval $(x_{i-1},x_i)$ with the jump point, will exceed the maximum difference $\mu$, found over all intervals belonging to the domain of continuity of the function.
+This value is set by the researcher.
+In our experiments it was equal to 3.
 
-When at the current $k$-th iteration of the method there is no $p$ that satisfies these conditions, it is interpreted as the absence of unspecified discontinuities.
-If such $p$ exists, this situation corresponds to the presence of undefined discontinuities.
-In this case, zero differences $\mu_i$ from (\ref{mu_i}), corresponding to intervals containing specified discontinuity points, are excluded from consideration by the condition $p<m-2(s+1)$.
+The lack of $p$ at the $k$-th iteration of the method that satisfies these conditions is interpreted as the absence of unspecified discontinuities.
+If such $p$ exists, it indicates the existence of undefined discontinuities.
+In this case, zero differences $\mu_i$ from (\ref{mu_i}) corresponding to intervals containing specified discontinuity points, are excluded from consideration by the condition $p<m-2(s+1)$.
 
@@
 %Russian
-Значение параметра $q$ отражает априорные предполжения о том, какая доля поисковых интервалов в их общем числе не будет содержать точек локальных экстремумов.
+The parameter $0<q<1$ is also set by the researcher.
+The value $1 - q$ reflects a priori assumptions about what proportion of search intervals will have $\mu_i$ values close to zero, i.e.
+actually characterizes the degree of multi-extremity of the problem.
+In our experiments, we used the value $q=0.3$.
 
@@
 
-\section{Algorithm for unspecified discontinuities}
+\section{Algorithm for Unspecified Discontinuities}
 
-The algorithm is a modification of the algorithm described in Section 2 for given discontinuities and consists of the following.
+This algorithm is a modification of the algorithm described in Section 2 for given discontinuities and consists of the following.
 
@@
 \begin{enumerate}
-\item Sort points of a set $\{x^1,...,x^k\} \cup \Omega$ in ascending order of coordinates 
+\item Sort points of a set $\{x^1,...,x^k\} \cup \Omega$ in ascending order by their coordinates 
 \begin{equation}\label{x_cond}
@@
 where $m=s+k+1$.
-Each point is matched with an attribute $\gamma_i$ from (\ref{gamma}) (the presence or absence of a set discontinuity) and with the value of the function $z_i=\varphi(x_i)$ calculated only for $x_i \not\in \Omega$, i.e.
+Each point is matched with an attribute $\gamma_i$ from (\ref{gamma}) (the presence or absence of a given discontinuity) and with the value of the function $z_i=\varphi(x_i)$ calculated only for $x_i \not\in \Omega$, i.e.
 at $\gamma_i=0$.
@@
 
-\item Determine $p$ satisfying the inequalities (\ref{mu_main_cond}), where $0<q<1<Q$ are the parameters of the algorithm.
+\item Determine $p$ which satisfies the inequalities (\ref{mu_main_cond}), where $0<q<1<Q$ are the parameters of the algorithm.
 Construct a set $I$ from (\ref{set_I}), the elements of which are interpreted as numbers of intervals that can contain a discontinuity.
@@
 $ 0 \leq \gamma_i+\gamma_{i-1}+|\delta_i | \leq 1, \; 1 \leq i \leq m, $
-are met.
+are satisfied.
 
@@
 
-\item Check the termination condition $x_t-x_{t-1} \leq \epsilon$, where $t$ is the number of the interval with the maximal characteristic and $\epsilon > 0$ is the predefined accuracy.
+\item Check the stopping condition $x_t-x_{t-1} \leq \epsilon$, where $t$ is the number of the interval with the maximal characteristic and $\epsilon > 0$ is the predefined accuracy.
 \end{enumerate}
@@
 \textbf{Remark}.
-В соответствии с (\ref{gamma}) and (\ref{delta_cond}) значение $\gamma_i+\gamma_{i-1}+|\delta_i |$ может быть равно 0 либо 1, в зависимости от наличия либо отсутствия точки разрыва в интервале $(x_{i-1},x_i)$.
-Так, если данный интервал не содержит точек разрыва, т.е. $\gamma_i+\gamma_{i-1}+|\delta_i | = 0$, то его характеристика (\ref{R_i}) примет вид 
+According to (\ref{gamma}) and (\ref{delta_cond}) the value $\gamma_i+\gamma_{i-1}+|\delta_i |$ can be 0 or 1, depending on the presence or absence of a jump in the interval  $(x_{i-1},x_i)$.
+So, if the given interval does not contain jumps, i.e. $\gamma_i+\gamma_{i-1}+|\delta_i | = 0$, then its characteristic  (\ref{R_i}) takes the form
 \[
@@
 \]
-Если одна из граничных точек интервала является заданной точкой разрыва, т.е. $\gamma_{i-1} = 1$ или $\gamma_i = 1$, то мы получим соответственно характеристики
+If one of the boundary points of the interval is a given discontinuity, i.e. $\gamma_{i-1} = 1$ or $\gamma_i = 1$, then we get the characteristics
 \begin{equation}\label{RR}
@@
 \end{equation}
-Аналогичные формулы для характеристик интервалов будут получаться, если $(x_{i-1},x_i)$ является интервалом, который содержит незаданную точку разрыва, т.е.
-когда $|\delta_i | = 1 $.
-В этом случае мы таже будем получать формулы (\ref{RR}) для $\delta_i = -1$ и $\delta_i = 1$, что соответствует использованию результатов испытания в правой точке интервала для построения его характеристики в случае убывания функции и выбору левой точки при возрастании функции.
-
-Fig.~\ref{ris2} shows the results of solving the problem from the previous example (\ref{testFunction}), in which the discontinuity points were considered unspecified.
-Here we used the parameters of the method $r=2.0, \epsilon=0.001, Q=3.0, q=0.3$.
-The dashes under the graph indicate the points of 70 trials required for the algorithm to solve problems with the specified accuracy.
+Similar formulas for the characteristics of intervals will be obtained if $(x_{i-1},x_i)$ is an interval that contains an unspecified discontinuity, i.e.
+when $|\delta_i | = 1 $.
+In this case, we will also obtain formulas (\ref{RR}) for $\delta_i = -1$ and $\delta_i = 1$, which correspond to using the trial result at the right point of the interval to compute its characteristic in the case of decreasing function and choosing left point as the function increases.
 
@@
 
+Fig.~\ref{ris2} shows the results of solving the problem from the previous example (\ref{testFunction}), in which the discontinuity points were unspecified.
+Here we used the parameters of the method $r=2.0, \epsilon=0.001, Q=3.0, q=0.3$.
+The dashes under the graph indicate the points of 70 trials required for the algorithm to solve problems with the specified accuracy.
+
 \begin{figure}%[!ht]
@@
 
-The proposed algorithm also allows partial specification of known discontinuities.
+The proposed algorithm also allows for partial specification of known discontinuities.
 Fig.~\ref{ris3} shows the results of solving the same problem, for which the discontinuities were considered partially given (the discontinuities at the points 4.6 and 9.0 were considered to be known).
@@
 
-To demonstrate the efficiency of the global search algorithm for discontinuous functions (GSA-D) described in the previous section, we compare it with Direct Search (DS) \cite{Audet}, Simulated Annealing (SA) \cite{Kirkpatrick}, and Genetic Algorithm (GA) \cite{Goldberg} implemented in MATLAB Global Optimization Toolbox \cite{MatlabOTB}.
+To demonstrate the efficiency of the global search algorithm for discontinuous functions (GSA-D) described in the previous section, we compared it with Direct Search (DS) \cite{Audet}, Simulated Annealing (SA) \cite{Kirkpatrick}, and Genetic Algorithm (GA) \cite{Goldberg} implemented in MATLAB Global Optimization Toolbox \cite{MatlabOTB}.
 
-As the main comparison criterion, we will use the number of search trials $K$ (i.e., the number of calculations of the objective function) performed by the method before it converges.
+We  used the number of search trials $K$ (i.e., the number of calculations of the objective function) performed by the method before it converges as the main comparison criterion.
 
-The algorithms were compared when solving 1000 problems with a discontinuous objective function, which were constructed as follows. First, we generated continuous functions of the form
+The algorithms were compared on 1000 problems with a discontinuous objective function, which were constructed as follows. First, we generated continuous functions of the form
 \begin{equation}\label{seriaFunction}
@@
 
-Since the solution of the test problem assumes that the global optimizer $x^*$ is known, its value was preemptively estimated for each function of the series by iterating through all nodes of a uniform grid, the step of which was chosen to be sufficiently small.
+Since the solution of the test problem assumes that the global optimizer $x^*$ is known, its value was preemptively estimated for each function of the series by iterating through all nodes of a uniform grid with the sufficiently small step.
 
-The first experiment was performed when solving a series of 1000 problems with continuous functions of the form (\ref{seriaFunction}).
-Table~\ref{tab1} shows the number of problems solved, the average number of iterations performed by different methods for solving the series, and the number of trials performed by the methods during the search.
-The problem was considered solved if the condition $|x^k-x^* | \leq \delta$ was met for the trial point $x^k$, where $x^*$ is the known solution of the problem and $\delta= 10^{-2}$.
+The first experiment was performed on a series of 1000 problems with continuous functions of the form (\ref{seriaFunction}).
+Table~\ref{tab1} shows the number of problems solved, the average number of iterations performed by different methods, and the number of trials performed by the methods during the search.
+The problem was considered solved if the condition $|x^k-x^* | \leq \delta$ was satisfied for the trial point $x^k$, where $x^*$ is the known solution of the problem and $\delta= 10^{-2}$.
 At the same time, the accuracy used in the termination criteria for all methods was $\epsilon = 10^{-3}$.
 
-In order to exclude the influence of inherent randomness when running algorithms from MatLab, the GA, SA, and DS methods were run 5 times for each problem, and the best result was recorded (in terms of the proximity of the found solution to the true one).
+In order to exclude the influence onherent randomness when running algorithms from MatLab, the GA, SA, and DS methods were run 5 times for each problem, and only the best result was recorded (in terms of the proximity of the found solution to the true one).
 In this case, the GA and SA methods were run with default parameters, while in the DS method the parameter $x_0$ (the starting point of the search) varied from 0 to 1 with a step of 0.2.
 
-With multiple runs, the methods from MATLAB Global Optimization Toolbox solved almost all the problems in the series, while with a single run of these methods the number of correctly solved problems was almost 2 times less.
+With multiple runs, the methods from MATLAB Global Optimization Toolbox solved almost all the problems in the series, while with a single run of these methods the number of correctly solved problems was almost 2 times lower.
 At the same time, the GSA-D method successfully solved all the problems in the series (the reliability parameter $r=2.2$ was used for the method).
@@
 
-The next experiment was performed when solving a series of 1000 problems with discontinuous functions of the form (\ref{seriaFunction_jump}).
+The next experiment was performed on a series of 1000 problems with discontinuous functions of the form (\ref{seriaFunction_jump}).
 GSA-D used the parameters $r=3.7, \; Q=3.0, \; q=0.3$, the other experimental conditions were similar to the previous run.
@@
 
-The results show that the Global Optimization Toolbox methods solve discontinuous problems somewhat worse than continuous ones, while the GSA-D algorithm proposed in the paper successfully solves the entire series of problems.
-The GA and SA algorithms show high reliability of finding the global optimizer after several runs, but these methods significantly lose to the GSA-D algorithm in terms of the number of objective function calculations.
-The DS method, showing a comparable number of trials, is inferior in the number of correctly solved problems.
+The results show that the Global Optimization Toolbox methods are somewhat worse at solving discontinuous problems than continuous ones, while the GSA-D algorithm proposed in the paper successfully solves the entire series of problems.
+The GA and SA algorithms show high reliability of finding the global optimizer after several runs, but these methods are significantly inferior to the GSA-D algorithm in terms of the number of objective function calculations.
+The DS method, while being able to solve the problem in a comparable number of trials, is inferior in the number of correctly solved problems.
 
-Summing up, we note that although the methods from the MATLAB Global Optimization Toolbox are stated as methods that work with discontinuous functions, they do not guarantee convergence to the global minimum.
+Although the methods from the MATLAB Global Optimization Toolbox are claimed to work with discontinuous functions, they do not guarantee convergence to the global minimum.
 The convergence rate (in terms of the number of trials) for the GA and SA methods is significantly lower than that for the DS method.
-However, the DS algorithm, while surpassing the GA and SA methods in this indicator, has the lowest reliability among the considered methods.
-The GSA-D method proposed in the article demonstrated both the guaranteed convergence to the global minimizer for the problems with discontinuous functions and the smaller number of trials required to achieve convergence.
+However, the DS algorithm, while surpassing the GA and SA methods by this indicator, has the lowest reliability among the considered methods.
+The GSA-D method proposed in the article demonstrated both the guaranteed convergence to the global minimizer for the problems with discontinuous functions and uses less  trials to achieve convergence.
 
@@
 In this paper, we propose an algorithm for finding the minimum of a multiextremal function with jump discontinuities.
-The proposed GSA-D algorithm, unlike many known approaches, does not use the concept of a gradient and provides a search for a global solution to the problem.
+The proposed GSA-D algorithm, unlike many known approaches, does not rely on the concept of a gradient and provides a search for a global solution to the problem.
 We carried out computational experiments confirming the convergence of the algorithm for a wide class of discontinuous multiextremal problems.
-The convergence conditions of the proposed algorithm, in general, will correspond to the convergence conditions of its prototype -- the global search algorithm for continuous problems, considered in detail in \cite{Strongin2000}.
+The convergence conditions of the proposed algorithm, in general, correspond to the convergence conditions of its prototype -- the global search algorithm for continuous problems, described in detail in \cite{Strongin2000}.
 Justification and research of the theoretical properties of the proposed GSA-D algorithm will be the subject of future publications.

### 2021-11-30 2022/PAVT/PAVT2022.tex
template for PAVT2022 paper

--- 2021-04-15 2021/RSD_Ufa/RSD2021.tex
@@
 \mainmatter              % start of a contribution
+% Решение обратных задач химической кинетки с помощью асинхронного алгоритма глобального поиска
+\title{Solving the Inverse Problems of Chemical Kinetics Using the Asynchronous Global Optimization Algorithm}
 %
-\title{Parallel Global Search Algorithm for Optimization of the Kinetic Parameters of Chemical Reactions}
-%
-\titlerunning{Parallel Global Search}  % abbreviated title (for running head)
+\titlerunning{Solving the Inverse Problems of Chemical Kinetics}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 \and
-Konstantin Barkalov$^4$\orcidID{0000-0001-5273-2471}
+Konstantin Barkalov$^4$ \letter \orcidID{0000-0001-5273-2471}
 \and
@@
 \begin{abstract}
-The paper considers the application of parallel computing technology to the simulation of a catalytic chemical reaction, which is widely used in the modern chemical industry to produce synthesis gas.
-As a chemical reaction, the process of pre-reforming propane on a Ni catalyst is assumed.
-To simulate a chemical process, it is necessary to develop a kinetic model of the process, that is, to determine the kinetic parameters.
-To do this, the inverse problem of chemical kinetics is solved, which predicts the values of kinetic parameters based on laboratory data.
-From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem.
-A parallel information-statistical global search algorithm was used to solve it.
-The use of the parallel algorithm has significantly reduced the search time to find the optimum.
-The found optimal parameters of the model made it possible to adequately simulate the process of pre-reforming propane on a Ni-catalyst.
 
+%обновить ключевые слова
 \keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Chemical kinetics $\cdot$ Inverse problems }
@@
 
-%Одним из направлений лаборатории математической химии Института нефтехимии и катализа УФИЦ РАН является исследование химических реакций методами математического моделирования.
-На данный момент изучено большое количество сложных химических процессов -- реакция циклоалюминирования алкенов триэтилалюминием в алюмациклопентаны, реакция гидроалюминирования олефинов алюминийорганическими соединениями, реакция спиртов с диметилкарбонатом и др.
-Математическое описание химических реакций необходимо для построения математических моделей химической кинетики, которые называются кинетическими моделями.
-Кинетическая модель является основой математического моделирования химических реакций и исходным уровнем при математическом моделировании химических реакторов.
-При разработке кинетической модели многостадийной химической реакции решается обратная задача химической кинетики (ХК).
-Обратной задачей ХК в общем случае называют процесс выбора механизма сложного процесса и определение его количественных характеристик -- констант скорости одностадийных реакций на основе опытных данных о динамике протекания процесса.
-С математической точки зрения обратная задача химической кинетики представляет собой задачу оптимизации, поэтому актуальной является задача выбора метода оптимизации.
-
-One of the areas of the Laboratory of Mathematical Chemistry of the Institute of Petrochemistry and Catalysis of the UFIC RAS is the study of chemical reactions by mathematical modeling methods.
-At the moment, a large number of complex chemical processes have been studied, among  them, the reaction of cycloalumination of alkenes with triethylaluminium into alumacyclopentanes, the reaction of hydroalumination of olefins with organoaluminum compounds, the reaction of alcohols with dimethyl carbonate, etc.
-The mathematical description of chemical reactions is necessary for the development the mathematical models of chemical kinetics, which are called kinetic models.
-The kinetic model is the basis for mathematical modeling of chemical reactions and the initial level for mathematical modeling of chemical reactors.
-When developing a kinetic model of a multi-stage chemical reaction, the inverse problem of chemical kinetics (CK) is solved.
-The inverse problem of CK is generally called the process of selecting the mechanism of a complex process and determining its quantitative characteristics -- the rate constants of reactions based on experimental data about the dynamics of the process.
-From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem, so the problem of choosing an optimization method is relevant.
-
-%Метод решения оптимизационных задач подобного типа должен учитывать следующую специфику.
-Во-первых, целевая функция задачи является "черным ящиком", т.к.
-она задается не в виде формулы, а в виде некоторого алгоритма вычисления ее значений.
-Во-вторых, вычисление даже одного значения целевой функции является трудоемкой операцией, т.к.
-требует численного моделирования.
-Перечисленная специфика не позволяет задействовать для решения обратных задач мультистартовые методы, основанные на идеях многократного градиентного спуска.
-С одной стороны, вычисление градиента в данном случае будет чрезмерно дорогой операцией, а с другой стороны - мультистартовые методы слабо применимы для существенно многоэкстремальных задач, какими и являются обратные задачи химической кинетики.
-Перечисленные особенности inverse problem of chemical kinetics учитываются в разработанном в University of Nizhni Novgorod параллельном алгоритме глобальной оптимизации.
-Теоретическим базисом данного алгоритма является information-statistical approach, детально изложенный в \cite{Strongin2000}.
-Разработанный алгоритм основан на редукции исходной многомерной задачи к эквивалентной ей одномерной задаче с последующим ее решением эффективными методами оптимизации функций одной переменной.
-Вопросы распараллеливания алгоритма на разных архитектурах рассмотрены в \cite{Barkalov2016}.
-В данной статье представлены результаты, полученные при использовании parallel global search algorithm для решения inverse problem of chemical kinetics.
-
-A method for solving the optimization problems of such type should account for the following aspects.
-First, the objective function of the problem is a ``black box'' since it is defined not by a formula but in the form of some algorithm for computing its values.
-Second, computing even a single value of the objective function is time-consuming operation since it requires numerical modeling.
-The aspects listed above don't allow employing the multistart methods based on the ideas of multiple gradient descents for solving the inverse problems.
-On one hand, computing the gradients would be too expensive operation in this case.
-On the other hand, the multistart methods are hardly applicable for the essentially multiextremal problems, which the inverse problems of chemical kinetics belong to.
-
-The features of inverse problems of chemical kinetics listed above were taken into account in the parallel global search algorithm developed in University of Nizhni Novgorod.
-The information-statistical approach described in \cite{Strongin2000} in details makes a theoretical basis for this algorithm.
-The developed algorithm is based on a reduction of the initial multidimensional problem to an equivalent one-dimensional problem followed by solving the latter by efficient optimization methods for the functions of single variable.
-The issues of the parallelization of the algorithm on various architectures were considered in \cite{Barkalov2016,Strongin2018,Barkalov2020}.
-In the present paper, the results obtained with the use of the parallel global search algorithm for solving the inverse problem of chemical kinetics are presented.
-
 \section{Mathematical Model}\label{Sec_math_mod}
-
-The process under study is propane pre-reforming into methane-rich gas over Ni catalyst, which is an industrially important chemical process \cite{USKOV2019126741, Uskov_en}.
-Pre-reforming of propane was studied over industrial nickel-chromium catalyst at pressure of 1 and 5 bar in the temperature range of 220 -- 380 $^0$C and at flow rates of 4000 and 12000 h$^{-1}$.
-The experimental data on propane pre-reforming was acquired from our previous work \cite{Uskov_en}.
-The reaction scheme consists of two reactions: propane steam conversion and CO$_2$ methanation \cite{Uskov_cat}:
-\begin{gather}
-C_3H_8 + 6H_2O \rightarrow 10H_2 + 3CO_2 \label{reac_ref}\\
-CO_2 + 4H_2 \rightleftharpoons CH_4 + 2H_2O \label{reac_met},
-\end{gather}
-where the rates of reactions \eqref{reac_ref} -- \eqref{reac_met} are expressed according to the Langmuir-Hinshelwood model \cite{USKOV2019126741}:
-
-\[
-W_{ref} = \dfrac{k_{ref}^0 \cdot \exp \left(- \dfrac{E_{ref}}{RT} \right) }{\left(1 + B \cdot C_{C3H8}\right)^m},
-\]
-
-\[
-W_{met} = k_{met}^0 \cdot \exp \left(- \dfrac{E_{met}}{RT} \right) \cdot C_{H2} \cdot \left[1 - \dfrac{P_{CH4} P_{H2O}^2}{K_{eq} P_{CO2} P_{H2}^4 }\right] ,
-\]
-where $W_{ref}$ and $W_{met}$ are the reaction rates; $E_{ref}$ and $E_{met}$ are the observed activation energies, J/mol; $k^0_{ref}$ and $k^0_{met}$ are the pre–exponential multipliers; $B$ is the constant parameter, $T$ is the temperature, K; $R$ is the universal gas constant, J/(K$\cdot$ mol).
-The ``$ref$''and ``$met$'' indexes refer to pre-reforming and methanation reactions, respectively.
-$C_{C3H8}$ and $C_{H2}$ are concentrations of propane and hydrogen, mol/m$^3$; $m$ is order of the denominator, which varied from 0 to 2; $K_{eq}$ is the equilibrium constant of $CO_2$ methanation; $P_{CH4}$, $P_{H2O}$, $P_{CO2}$, $P_{H2}$ are partial pressures of the corresponding substances, bar.
-The mathematical model is a system of equations of material balance:
-
-\begin{equation}\label{mater_balance}
-\begin{cases}
-G \dfrac{dy_i}{dl} = \left(\nu_i^{ref} W_{ref} + \nu_i^{met} W_{met} \right) m_i, 
-\\
-0 \le l \le , i \in \left\lbrace C_3H_8, CH_4, H_2O, H_2, CO_2 \right\rbrace,
-\\
-l = 0, y_i = y_{i0},
-\end{cases}
-\end{equation}
-where $G$ is a mass flow of the mixture, kg/(m$^2 \cdot$ sec); $y_i$ is a mass fraction of the $i$-th component; $\nu_i$ is a stoichiometric coefficient of the $i$-th component; $m_i$ is a molar mass of the $i$-th component, kg/mol; $l$ is coordinate along the catalytic layer, m; $L$ is a length of the catalytic layer, m.
-The length of the catalytic layer is 0.008 m.
-The mathematical model of chemical kinetics problems is a system of differential equations that describes the variations in substance concentrations over time according to the rates of reaction stages.
-The system of differential equations is a Cauchy problem containing the initial data \cite{ENIKEEV2020123131, Chainikova}.
-The numerical solving of such a system of equations is a direct problem of chemical kinetics.
-
-%Системы обыкновенных дифференциальных уравнений (ОДУ), описывающие химические прцоессы, зачастуя являются жесткими за счет наличия быстропротекающих и медленнопротекающих стадий реакции.
-Поэтому в качестве метода решения системы ОДУ был выбран метод RADAU-IIA, который подходит для решения жестких систем ОДУ.
-Рассмотрим в общем случае задачу Коши вида:
-
-Systems of ordinary differential equations (ODES) describing chemical processes are often stiff due to the presence of fast-flowing and slow-flowing reactions.
-Therefore, the RADAU-IIA method was chosen as a method for solving the ODE system, which is suitable for solving stiff ODE systems.
-Let us consider in the general case the Cauchy problem of the form:
-
-\begin{equation}\label{ode}
-\dfrac{dx}{dt} = f(\textbf{x}, y), \textbf{x}(0) = \textbf{x}_0, 0 \le t \le T,
-\end{equation}
-%где x в нашем случае это вектор концентраций веществ реакции размерности 5, а f это правая часть уравнения \eqref{mater_balance}.
-Тогда s-стадийный метод Рунге-Кутты для решения системы \eqref{ode} выписывается следующим образом:
-where $x$ is the vector of concentrations of reaction substances with dimension of 5 (the substances are $C_3H_8$, $CH_4$, $H_2O$, $H_2$, $CO_2$), and $f$ is the right-hand side of the equation \eqref{mater_balance}.
-Then the $s$-stage Runge-Kutta method for solving the \eqref{ode} system is written as follows:
-
-\begin{equation}\label{runge}
-\begin{cases}
-X_i = x_n + h \sum\limits_{j=1}^s a_{ij} f(X_j, t_n + c_j h) , 1 \le i \le s, 
-\\
-x_{n+1} = x_n + h \sum\limits_{i=1}^s b_i f (X_i, t_n + c_i h),
-\end{cases}
-\end{equation}
-%где $x_n$, $x_{n+1}$, $X_i$ -- приближенные решения системы \eqref{ode} в моменты времени $t_n$, $t_{n+1} = t_n + h$, $t_n + c_i h$.
-Коэффициенты  методов  Рунге-Кутты  обычно  записывают  в  виде таблицы Бутчера:
-where $x_n$, $x_{n+1}$, $X_i$ are approximate solutions of the system \eqref{ode} at time points $t_n$, $t_{n+1} = t_n + h$, $t_n + c_i h$.
-The coefficients of the Runge-Kutta methods are usually written in the form of a Butcher tableau:
-
-\begin{equation}\label{butcher}
-\begin{array}{c|cccc}
-c_1 & a_{11} & a_{12} & ... & a_{1s} \\
-c_2 & a_{21} & a_{22} & ... & a_{2s} \\
-... & ... & ... & ... & ... \\
-c_s & a_{s1} & a_{s2} & ... & a_{ss} \\
-\hline
- & b_{1} & b_{2} & ... & b_{s} \\
-\end{array}
-\end{equation}
-
-% Таблица  Бутчера  содержит  векторы  b  и  с,  состоящие  из коэффициентов  bi  и  ci  соответственно  и  матрицу  A  размерностью $s \times s$, включающую  коэффициенты  aij. 
-%Метод Радо-IIA пятого порядка (Radau 5):
-
-The Butcher tableau contains the vectors $b$ and $c$, consisting of the coefficients $b_i$ and $c_i$, respectively, and the matrix $A$ of dimension $s \times s$, including the coefficients $a_{ij}$.
-The coefficients of fifth-order Rado-IIA method (Radau 5) are presented below:
-
-\begin{equation}\label{radau}
-\begin{array}{c|ccc}
-\dfrac{4 - \sqrt{6}}{10} & \dfrac{88 - 7\sqrt{6}}{360} & \dfrac{296 - 169\sqrt{6}}{1800} & \dfrac{-2 + 3\sqrt{6}}{225} \\
-\dfrac{4 + \sqrt{6}}{10} & \dfrac{296 + 169\sqrt{6}}{1800} & \dfrac{88 + 7\sqrt{6}}{360} & \dfrac{-2 - 3\sqrt{6}}{225} \\
-1 & \dfrac{16 - \sqrt{6}}{36} & \dfrac{16 + \sqrt{6}}{36} & \dfrac{1}{9} \\
-\hline
-& \dfrac{16 - \sqrt{6}}{36} & \dfrac{16 + \sqrt{6}}{36}  & \dfrac{1}{9} \\
-\end{array}
-\end{equation}
-
-The solution of the system of ordinary differential equations is a vector of the calculated concentrations of the reaction components, which are then compared with the experimental data.
-Determining the kinetic parameters of reaction stages by comparing calculated values of substance concentrations and experimental results is an inverse problem of chemical kinetics.
-The mathematical problem is to minimize the functional of the deviation between calculated and experimental values.
-The objective function determined as the sum of absolute deviations between calculated and experimental concentrations:
-
-\begin{equation}\label{func}
-F = \sum\limits_{i=1}^M \sum\limits_{j=1}^N \left| x_{ij}^{calc} - x_{ij}^{exp} \right| \rightarrow \min,
-\end{equation}
-where $x_{ij}^{calc}$ and $x_{ij}^{exp}$ are calculated and experimental values of component concentrations; $M$ is the number of measuring points; $N$ is the number of substances involved in the reaction.
 
@@
 
-Let us consider global optimization problem of the form 
-\begin{gather}
- \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problem}\\
- D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \; a_i,b_i\in R, \;  1\leq i \leq N\right\} \label{D},
-\end{gather}
-where the objective function is a black-box function and it is assumed to satisfy the Lipschitz condition
+Мастер-рабочие, мастер выполняет алгоритм и распределяет задания между рабочими.
+
+На начальной фазе поиска процесс-мастер инициирует параллельное проведение $p$ испытаний в $p$ произвольных внутренних точках области поиска $\{y(x^1), y(x^2), ...,y(x^p)\}$.
+
+Пусть выполнено $k$ испытаний (в частности, $k$ может быть равно 0), и проводятся испытания в точках $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$
+
+На основной фазе поиска каждый процесс-рабочий, завершивший испытание в некоторой точке (будем считать, что это точка $y(x^{k+1})$), информирует об этом процесс-мастер, который, в свою очередь, немедленно выбирает точку нового испытания $x^{k+p+1}$ в соответствии с правилами, описанными ниже.
+
+В данном случае мы будем иметь множество 
 \[
-\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D,
+I_k = \{x^{k+1},x^{k+2},...,x^{k+p}\}
 \]
-with the constant $L, \; L<\infty,$ unknown a priori.
+точек, в которых испытания уже начались, но еще не завершены.
 
-The assumption of the objective function to be Lipschitzian is typical of many approaches to the development of the deterministic global optimization algorithms.
-The first methods of Lipschitz optimization were proposed in the early 1970s \cite{Piyavskii1972,Shubert1972}.
-Since that time, this line of research has continued to develop actively \cite{Evtushenko2013,Zilinskas2010,Pinter1996,Jones2009}.
+Step 1. Точки множества 
+\[
+X_k
+\]
+перенумеровать в порядке возрастания координаты, т.е.
+\[
+0=
+\]
+Step 2. Вычислить значения 
+\[
+M_1=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin I_k, x_i \notin I_k, 2\leq i\leq k+p \right\},
+\]
+\[
+M_2=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in I_k, 2\leq i < k+p \right\},
+\]
+\[
+M=\max\{M_1,M_2\},
+\]
+где $z_i=\varphi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$. Значения $z_i$ в точках $x_i \in I_k$ являются неопределенными, т.к. испытания в точках $x_i \in I_k$ еще не завершены.
 
-Currently, nature-inspired algorithms are widely used for solving optimization problems with the black-box objective functions; for example, see \cite{Yang2013,Gendreau2010,Eiben2015}.
-Such algorithms, in one way or another, employ the ideas of random search.
-Due to simple implementation and use, they have become very popular.
-However, these algorithms are inferior to the deterministic counterparts in terms of quality \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a certain set).
+Написать про 0 или неопределеность
 
-One of the efficient deterministic methods for solving multiextremal optimization problems is \textit{the information-statistical global search algorithm} \cite{Strongin2000}.
-This method initially proposed for solving unconstrained optimization problems was successfully generalized to the classes of optimization problems with non-convex constraints and multicriteria optimization problems.
-For different versions of the algorithm, parallelization methods taking into account the architecture of modern computing systems were also suggested \cite{Barkalov2016,globalizerSystem,Strongin2018}.%поставить ссылку
-
-The parallel global search algorithm described in this section is based on the idea of reducing the dimension with the use of the Peano curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously map the unit interval $[0,1]$ onto the $N$-dimensional cube $D$ from (\ref{D}).
-By using this kind of mapping, it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem
-\[
-\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}},
-\]
-where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
-\[
-\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}
-\]
-with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation
-$ H=2 L \sqrt{N+3}$ and $y(x)$ is a Peano curve from $[0,1]$ onto $D$.
-
-Note that theoretically the Peano curve $y(x)$ is defined as a limit object.
-Therefore, in practical implementation, only some approximation to the true space-filling curve can be constructed.
-Some methods for constructing this type of approximations (called \textit{evolvents}) are considered in \cite{Sergeyev2013,Strongin2000}.
-In this case, the accuracy of the approximation to the true curve $y(x)$ depends on the density of the evolvent $m$ (which is a parameter for constructing the evolvent) and is of the order of $2^{-m}$ for each coordinate.
-Examples of evolvents with different values of $m$  for the dimensionality $N=3$ are shown in Fig.~\ref{evolvents}.
-
-\begin{figure}[ht]
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{fig1a.JPG} \\ (a)}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (b)}
-\end{minipage}
-\caption{Evolvents in three dimensions with (a) $M=3$ and (b) $M=4$}
-\label{evolvents}
-\end{figure}
-
-Let us call the process of computing a function value (including the construction of the image $y=y(x)$) as a \textit{trial}, and the pair $\{x, z = \varphi(y(x))\}$ as the outcome of the trial.
-When describing a parallel algorithm, we will use the term \textit {iteration} to denote the simultaneous (parallel) execution of several trials (each trial is executed by a separate processor).
-The number of trials during the $n$-th iteration we will denote as $p$ , and the total number of trials executed during all $n$ iterations as $k(n)$ .
-In other words, we assume that we have  $p$  processors at our disposal while executing the $n$-th iteration.
-
-The parallel global search algorithm used in this research (according to \cite{Strongin2000}) can be formulated as follows.
-The first $p$ trials are executed at the points $x^0 = 0$, $x^1 = 1$ and at the arbitrary internal points $x^2, ..., x^{p-1}$ of the interval $(0,1)$.
-Let us assume $n \geq 1$  iterations of the method to be completed, in the course of which the trials in $k=k(n)$ points $x^i, 0 \leq i \leq k,$ have been executed.
-Then, the points $x^{k+1},...,x^{k+p}$  of the search trials for the next $(n+1)$-th iteration are determined according to the following rules.
-
-\begin{enumerate}
-	\item 
-	Renumber the inverse images of all the points from the trials already performed  
-\begin{equation}\label{y_i}
-y^0=y(x^0), y^1=y(x^1),...,y^k=y(x^k)
-\end{equation}
-by subscripts in the increasing order of their coordinates, i.e.
-\begin{equation}\label{x_i}
-0=x_0<x_1<\dots <x_k=1,
-\end{equation}
-and associate these with the values $z_i=\varphi(y(x_i)), 0\leq i \leq k,$  computed at these points.
-\item
-Compute the maximum absolute value of the first divided differences
-\begin{equation}\label{mu}
-\mu = \max_{1 \leq i \leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
-\end{equation}
-where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$. If $\mu = 0$, set $\mu = 1$.
-\item
-For each interval $(x_{i-1}, x_i), \; 1\leq i \leq k,$  calculate the value
-\begin{equation}\label{R}
-R(i)=r\mu\Delta_i+\frac{(z_i-z_{i-1})^2}{r\mu\Delta_i}-2(z_i+z_{i-1})
-\end{equation}
-called the \textit{characteristic} of the interval; the real number $r>1$ being the input parameter of the algorithm.
-
-\item 
-Arrange characteristics  $R(i), 1 \leq i \leq k$, in decreasing order 
-\begin{equation}\label{Rdec}
-R(t_1)\geq R(t_2)\geq \dots \geq R(t_{k})
-\end{equation}
-and select $p$ largest characteristics with interval numbers $t_j, 1\leq j \leq p$.
-\item
-Carry out new trials at points $x^{k+j}\in(x_{t_j-1},x_{t_j}), 1\leq j\leq p$, computed according to the formula
-\[
-x^{k+j} = \frac{x_{t_j}+x_{t_j-1}}{2} - \mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N.
-\]
-
-\end{enumerate}
-
-The algorithm stops if condition $\Delta_{t_j}<\epsilon$ is satisfied for at least one index $t_j, 1 \leq j \leq p$ ; here $\epsilon>0$ is the predefined accuracy.
-Also, the algorithm stops if the maximum allowed number of trials $K_{max}$ preset prior to running is exhausted.
-
-%Russian
-The considered variant of the parallelization is \textit{synchronous}, when the transition to the next iteration is performed after full termination of the current one, i.e., after completing the last trial of the current iteration.
-Particular implementation of this parallelization scheme will depend on the computer system architecture as well as on the requirements to the software necessary for computing the objective function values.
-In the present study, the conducting of trials required solving a stiff ODE system (for solving this one, Radau IIA method was used).
-Therefore, the parallelization was organized on CPU using MPI technology.
-The global search algorithm was run on a single (master) processor.
-The rest (slave) processors conducted parallel trials within the Step 5 of the algorithm.
-The amount of data transferred between the processes was small: it was only required to send out the coordinates of trial points and then to collect the computed values of the objective function at these points.
-At the same time, the time of conducting a single trial was not less 1 sec that considerably exceeded the time of data transfer between the processes.
-This way, the overhead costs for the data transfer didn't essentially affect the parallelization efficiency.
-
-%Конкретная реализация данной схемы распараллеливания будет зависеть от архитектуры вычислительной системы, а также от требований к программному обеспечению, необходимому для вычисления значений целевой функции задачи.
-В данном исследовании проведение испытаний требовало решения жесткой системы ODE (для ее решения был использован метод Radau IIA).
-Поэтому распараллеливание было организовано на CPU с использованием технологии MPI.
-В одном (master) процессе работал алгоритм глобального поиска; остальные (slave) процессы проводили параллельные испытания в рамках выполнения шага 5 алгоритма.
-Объем передаваемых данных между процессами был небольшим: требовалось лишь разослать координаты точек испытаний и затем собрать вычисленные значения целевой функции в этих точках.
-Одновременно с этим время проведения одного испытания составляло не менее 1 сек., что значительно превышает время передачи данных между процессами.
-Тем самым, накладные расходы на передачу данных не оказывали существенного влияния на эффективность распараллеливания.
+Step 3. Каждому интервалу поставить в соответствие число
 
 \section{Numerical Experiments}\label{Sec_Exp}
-%Таким образом, в результате решения обратной задачи химической кинетики параллельным методом глобального поиска были рассчитаны кинетические параметры процесса предриформинга пропана на Ni катализаторе (Table \ref{tab:1}).
-Для интерпретации полученных результатов проводится решение прямой задачи химической кинетики, то есть расчет концентраций компонентов реакции и сравнение их с экспериментальными данными.
-На рисунках НОМЕР a, b показано сравнение рассчитанных концентраций с экспериментальными при разных условиях: (a) GHSV 4000 h$^{-1}$, (b) GHSV 12000 h$^{-1}$.
-Результаты расчетов качественно согласуются с экспериментом.
-Ошибка отклонения составляет не более ПОСЧИТАТЬ \%.
-Thus, as a result of solving the inverse problem of chemical kinetics by the parallel global search method, the kinetic parameters of the propane pre-reforming process on a Ni catalyst were calculated (Table \ref{tab:1}).
-To interpret the results obtained, the solution of the direct problem of chemical kinetics was carried out, namely, the concentrations of the reaction components were calculated and compared with the experimental data.
-Fig.~\ref{fig_kin_curves} shows a comparison of the calculated concentrations with the experimental ones under different conditions: (a) GHSV=4000 h$^{-1}$, (b) GHSV=12000 h$^{-1}$.
-The results of the calculations are qualitatively consistent with the experiment.
-
-\begin{figure}[ht]
-	\begin{minipage}{0.5\linewidth}
-		\center{\includegraphics[width=1.0\linewidth]{kin_curve1.png} \\ (a)}
-	\end{minipage}
-	\begin{minipage}{0.5\linewidth}
-		\center{\includegraphics[width=1.0\linewidth]{kin_curve2.png} \\ (b)}
-	\end{minipage}
-	\caption{Temperature dependences of the output concentrations of propane C3H8, methane CH4, hydrogen H2 and CO2 in the process of propane pre-reforming.
-Experimental conditions: 220-380 $^0$C, (a) GHSV = 4000 h$^{-1}$, (b) GHSV = 12000 h$^{-1}$, 1 bar pressure.
-Concentrations of the gas components on the figure are given on the dry basis.
-Points are experimental concentrations (``exp''-index), lines are simulated concentrations (``calc''-index).}
-	\label{fig_kin_curves}
-\end{figure}
-
-\begin{table}[ht]
-	\caption{Calculated kinetic parameters}
-	\label{tab:1}
-	\center
-	\begin{tabular}{c|c|c|c|c|c}
-		\hline
-		%\noalign{\smallskip}
-		$E_{ref}$, kJ/mol & $E_{met}$, kJ/mol & $k_{ref}$, (mole$\cdot$m$^{-3}$)m$\cdot$s$^{-1}$ & $k_{met}$, s$^{-1}$ & $m$ & $B$ \\
-		%\noalign{\smallskip} 
-		\hline 
-		%\noalign{\smallskip}
-		99.23 \;&	37.08 \; & $2.4 \cdot 10^{10}$ \;  & $1.93 \cdot 10^5$ \; & 0.82 & 4.8  \\
-		%\noalign{\smallskip}
-		\hline
-	\end{tabular}
-\end{table}
-%Для сравнения с предыдущими расчетами был построен график арренисовской зависимости (рисунок НОМЕР), на котором представлены зависимости десятичного логарифма от констант скоростей реакций \eqref{reac_ref} и \eqref{reac_met} от обратной температуры.
-Обозначения "GA", "GSA" и "PGSA" означают  решения, полученные различными алгоритмами оптимизации, соотвественно, генетическим алгоритмом, алгоритмом гравитационного поиска и параллельным алгоритмом глобального поиска.
-
-For comparison with the previous calculations, an Arrhenis plot was constructed (Fig.~\ref{fig_arren}), which shows the dependence of the decimal logarithm on the reaction rate constants \eqref{reac_ref} and \eqref{reac_met} on the inverse temperature.
-The designations ``GA'', ``GSA'' and ``PGSA'' mean solutions obtained by various optimization algorithms, the genetic algorithm, the gravity search algorithm, and the parallel global search algorithm, respectively.
-The blue lines refer to the pre-reforming reaction \eqref{reac_ref}, and the yellow lines refer to the methanation reaction~\eqref{reac_met}.
-Note that the lines of the same color are parallel to each other, which means that the activation energies of the reaction calculated by different methods have approximately the same values, which indicates the adequacy of the calculations.
-
-\begin{figure}[ht]
-	\center{\includegraphics[width=0.8\linewidth]{arren.png}}
-	\caption{Comparison of the obtained solution with the other algorithms found on the Arrhenius plot}
-	\label{fig_arren}
-\end{figure}
-
-Numerical experiments were performed on the Lobachevsky supercomputer of the University of Nizhni Novgorod (operating system -- CentOS 7.2, control system -- SLURM).
-One supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM.
-The CPU is 8-core (i.e.
-a total of 16 CPU cores are available on the node).
-The numerical experiments were performed using the Globalizer system \cite{globalizerSystem}.
-To build the system for running on the Lobachevsky supercomputer, the GCC 5.5.0 compiler and Intel MPI 2017 were used.
-To solve the stiff system of equations arising in computing the objective function values, Radau method from SciPy library from Python 3.9 was used.
-
-%Для решения жесткой системы уравнений, возникающей при вычислении значений целевой функции, использовался метод Radau из библиотеки SciPy, Python 3.9.
-Global search algorithm запускался со следующими параметрами: параметр надежности $r=5$, точность поиска $\epsilon = 0.01$, плотность построения evolvent $m=10$, максимальное число испытаний $K_{max}=5000$.
-После завершения работы алгоритма найденное решение уточнялось с помощью локального поиска, использовался Hooke--Jeeves method \cite{HookJeeves}.
-Точность локального поиска составляла $\epsilon = 0.001$.
-Для оценки эффективности распараллеливания global search algorithm запускался в последовательном и паралелльном режимах, максимальное число задействованных вычислительных узлов составляло 20, соответственно, максимальное число задействованных процессоров было 40.
-На Fig.~\ref{fig_time},\ref{fig_speedup} приведены графики, отражающие время и ускорение параллельного алгортима в зависимости от числа процессоров; время работы приведено в часах.
-
-Global search algorithm was run with the following parameters: the reliability parameter $r=5$, the search precision $\epsilon = 0.01$, the evolvent density $m=10$, the maximum allowed number of trials $K_{max}=5000$.
-Upon completing the execution of the algorithm, the solution found was refined with a local search using Hooke--Jeeves method \cite{HookJeeves}.
-The local search precision was $\epsilon = 0.001$.
-To evaluate the parallelization efficiency, the global search algorithm was run in the sequential regime and in the parallel one, the maximum number of computer nodes employed was 20.
-Correspondingly, the maximum number of processors employed was 40.
-In Fig.~\ref{fig_time} and Fig.~\ref{fig_speedup} the plots reflecting the running time and the speedup of the parallel algorithm {\em vs} the number of processors employed, respectively are presented (the time is in hours).
-
-\begin{figure}[ht]
-\center{\includegraphics[width=0.8\linewidth]{time.pdf}}
-\caption{The parallel algorithm running time}
-\label{fig_time}
-\end{figure}
-
-\begin{figure}[ht]
-\center{\includegraphics[width=0.8\linewidth]{speedup.pdf}}
-\caption{Speedup of the parallel algorithm}
-\label{fig_speedup}
-\end{figure}
 
 \section{Conclusions and Future Work}
-
-%В будущем планируется моделирование других химических процессов, а именно, первая задача это моделирование процесса сернокислотного алкилирования изобутана олефинами, для которого на данный момент составлена схема химических превращений, проведен литературный обзор процесса и найдены экспериментальные данные.
-Математическая модель реакции сернокислотного алкилирования изобутана олефинами представляет собой задачу Коши, на данный момент решена прямая задача химической кинетики, далее будет решаться обратная задача предложенными параллельными методами.
-Вторая задача это моделирование процесса синтеза метил-трет бутилового эфира.
-Химизм данного процесса известен, задача иемют небольшую размерность, однако моделирование реактора трудоемко и представляет собой решение уравнение в частных производных второго порядка.
-На данный момент расписана разностная схема, реализован метод прогонки, следующим этапом будет решение прямой задачи химической кинетики, а также предсказание кинетических параметров процесса предложенными параллельными методами.
-
-In the future, it is planned to simulate other chemical processes, namely, the first task is to simulate the process of sulfuric acid alkylation of isobutane with olefins, for which a scheme of chemical transformations has been compiled, a literature review of the process has been conducted, and experimental data have been found.
-The mathematical model of the reaction of sulfuric acid alkylation of isobutane by olefins is a Cauchy problem, at the moment the direct problem of chemical kinetics is solved, then the inverse problem will be solved by the proposed parallel methods.
-The second task is to simulate the synthesis of methyl-tert-butyl ether.
-The chemistry of this process is known, and the problem has a small dimension, but the simulation of the reactor is time-consuming and represents the solution of a second-order partial differential equation.
-At the moment, the difference scheme is formed, the run-through method is implemented, the next stage will be the solution of the direct problem of chemical kinetics, as well as the prediction of the kinetic parameters of the process by the proposed parallel methods.
-
-%В планы дальнейших работ входит также и повышение эффективности распараллеливания используемого метода.
-Относительно невысокие показатели ускорения при решении данной задачи объясняются двумя обстоятельствами.
-Во-первых, при реализации параллельного алгоритма была применена синхронная схема распараллеливания.
-Управляющий процесс (в котором работал метод оптимизации) при выполнении  шага 5 алгоритма ожидал окончания проведения всех испытаний, проводимых подчиненными процессами.
-Тогда как в рассматриваемой задаче время проведения одного испытания (как было экспериментально установлено) варьировалось от 1 до 40 секунд в зависимости от конкретных значений параметров.
-Возможный путь к повышению эффективности распаралелливания - реализация асинхронной схемы проведения испытаний на шаге 5 алгоритма.
-В этом случае результаты испытания, полученные в одном из подчиненных процессов, могут быть отправлены для обработки в управляющий процесс немедленно, без ожидания окончания других испытаний.
-После чего подчиненный процесс получает от управляющего процесса новую точку, в которой проводит очередное испытание.
-Асинхронная схема распараллеливания, теоретически, способна обеспечить полную загрузку задействованных вычислительных ресурсов даже в том случае, когда время выполнения испытаний различно в разных точках области поиска.
-Во-вторых, при масштабном распараллеливании "узким местом" становится локальный поиск, который выполняется для уточнения решения после завершения фазы глобального поиска.
-Локальный поиск требует незначительного (по сравнению с глобальным) числа trials, но эти испытания выполняются последовательно.
-И при паралелльном запуске это незначительное число последовательных испытаний становится сопоставимым с числом параллельных итераций глобального поиска.
-В дальнейшем планируется использовать параллельную реализацию метода локального поиска на завершающем этапе работы, что способно повысить эффективность распараллеливания в целом.
-
-The plans of further work include also the improvement of the parallelization efficiency for the global optimization method used.
-Relatively low indicators of speedup in the solving of the present problem can be explained by two factors.
-First, a synchronous parallelization scheme was applied in the implementation of the parallel algorithm.
-The master process (which the optimization method worked in) was waiting for the finish of execution of all trials performed by the slave processes when executing Step 5 of the algorithm.
-However, the time of executing  a single trial in the problem considered varied from 1 up to 40 seconds subject to particular values of parameters (as it was found experimentally).
-The implementation of the asynchronous scheme of executing the trials at Step 5 of the algorithm seems to be a promising way of improving the parallelization efficiency.
-In this case, the trial results obtained by a slave process may be sent to the master process for processing immediately without waiting for the competing of other trials.
-Afterwards, the slave process receives a new point of the next trial form the master one.
-Theoretically, the asynchronous parallelization scheme can provide a full load of the computer resources employed even if the trial execution times are different at different points of the search domain.
-
-Second, the local search, which is performed to refine the solution after completing the global search phase becomes a ``bottleneck'' at large-scale parallelization.
-The local search requires insufficient number of trials (as compared to the global one), but these trials should be performed sequentially.
-And at the parallel run this insufficient number of sequential trials becomes comparable to the number of the parallel global search iterations.
-In future, we plan to use the parallel implementation of the local search method in the final stage of the work that can improve the overall parallelization efficiency.
 

### 2021-12-07 2022/PAVT/PAVT2022.tex
UNN part in Russian

@@
 % Решение обратных задач химической кинетки с помощью асинхронного алгоритма глобального поиска
+% Решение обратных задач химической кинетки с помощью смешанного локально-глобального поискового алгоритма 
 \title{Solving the Inverse Problems of Chemical Kinetics Using the Asynchronous Global Optimization Algorithm}
@@
 
+The paper considers the application of parallel computing technology to the simulation of a catalytic chemical reaction, which is widely used in the modern chemical industry to produce synthesis gas.
+As a chemical reaction, the process of pre-reforming propane on a Ni catalyst is assumed.
+To simulate a chemical process, it is necessary to develop a kinetic model of the process, that is, to determine the kinetic parameters.
+To do this, the inverse problem of chemical kinetics is solved, which predicts the values of kinetic parameters based on laboratory data.
+From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem.
+A parallel information-statistical global search algorithm was used to solve it.
+The use of the parallel algorithm has significantly reduced the search time to find the optimum.
+The found optimal parameters of the model made it possible to adequately simulate the process of pre-reforming propane on a Ni-catalyst.
+
 %обновить ключевые слова
@@
 
-\section{Mathematical Model}\label{Sec_math_mod}
+%Часть УГНТУ
 
-\section{Parallel Global Search Algorithm}\label{Sec_GSA}
+%Стыковочный абзац
+Постоянно возрастающая сложность исследуемых реакций влечет за собой усложнение их математических моделей, что существенно затрудняет поиск кинетических констант.
+Как правило, найти кинетические константы реакций аналитическим путем не представляется возможным, поэтому возникает необходимость в разработке и использовании численных методов их поиска (см., например, [ссылки]).
+В этом случае критерий качества найденного решения (целевая функция) не имеет явного аналитического описания, но допускает алгоритмическое представление и требует значительных затрат на расчет.
+Кроме того, в обратных задачах химической кинетики целевая функция может быть существенно многоэкстремальной, т.е.
+иметь много локальных экстремумов наряду с глобальным.
 
-Мастер-рабочие, мастер выполняет алгоритм и распределяет задания между рабочими.
+%Часть ННГУ
+Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) существенно отличаются от методов локального поиска (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}).
+Как правило, методы локальной оптимизации не могут покинуть области притяжения локального экстремума и не находят глобальный оптимум.
+А использование параметров модели, соответствующих найденному локальному решению, может оказаться недостаточным, поскольку глобальное решение может дать существенный выигрыш по сравнению с локальным.
 
-На начальной фазе поиска процесс-мастер инициирует параллельное проведение $p$ испытаний в $p$ произвольных внутренних точках области поиска $\{y(x^1), y(x^2), ...,y(x^p)\}$.
+Разнообразие возникающих задач глобальной оптимизации влечет за собой разнообразие подходов к их решению.
+Методы решения задач глобальной оптмизации можно условно разделить на два класса: метаэвристические и детерминированные.
+Метаэвристические алгоритмы, как правило, основаны на имитации процессов, протекающих в живой природе.
+Примерами метаэвристических алгоритмов являются simulated annealing, evolution and genetic algorithms и т.д.
+(see, e.g., \cite{Battiti2009,Eiben2015}).
+В силу своей относительной простоты метаэвристические алгоритмы более популярны у исследователей, чем детерминированные.
+Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может распологаться далеко от глобального \cite{Kvasov2018}.
 
-Пусть выполнено $k$ испытаний (в частности, $k$ может быть равно 0), и проводятся испытания в точках $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$
+Возможность построения детерминированных методов глобального поиска, отличающихся от перебора по сетке и метаэвристических алгоритмов, связана с наличием и учетом  априорных предположений о свойствах функций задачи.
+Такие предположения играют ключевую роль при разработке эффективных алгоритмов глобальной оптимизации и служат основным математическим инструментом для получения оценок глобального решения.
 
-На основной фазе поиска каждый процесс-рабочий, завершивший испытание в некоторой точке (будем считать, что это точка $y(x^{k+1})$), информирует об этом процесс-мастер, который, в свою очередь, немедленно выбирает точку нового испытания $x^{k+p+1}$ в соответствии с правилами, описанными ниже.
+Одним из естественных допущений о задаче является предположение об ограниченности относительных изменений значений целевой функции.
+Такое допущение связано с тем, что отношения приращений функции к соответствующим приращениям ее аргумента обычно не могут превышать некоторого порога, определяемого ограниченной энергией изменений в моделируемой системе.
+В таком случае, как известно, функции называются липшицевыми, а сама задача -- задачей липшицевой глобальной оптимизации.
 
-В данном случае мы будем иметь множество 
+В данной статье отражены результаты применения параллельных методов липшицевой оптимизации, разрабатываемых в ННГУ им. Н.И. Лобачевского, к решению обратных задач химической кинетики. 
+Основная часть статьи имеет следующую структуру.
+Описание математической модели исследуемой химической реакции приведено в Section 2.
+Формальная постановка задачи липшицевой глобальной оптимизации приводятся в Section 3.
+В Section 4 кратко изложена схема разрабатываемого авторами асинхронного параллельного алгоритма для решения задач указанного класса.
+Результаты численного решения обратной задачи химической кинетики обсуждаются в Section 5.
+
+\section{Problem Statement}\label{Sec_math_mod}
+
+%Содержательная постановка задачи
+
+\section{}\label{Sec_GSA}
+
+\subsection{Global Optimization Problem}
+
+%Черновая заготовка для описания параллельного алгоритма оптимизации
+Как уже говорилось выше, с формальной точки зрения the inverse problem of chemical kinetics рассматривается нами как Lipschitz global optimization problem. 
+В общем виде задача указанного класса может быть математически сформулирована следующим образом:
+\begin{gather}
+ \varphi^* = \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problemN}\\
+ D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\} \label{D},
+\end{gather}
+где $a,\; b$ есть заданные векторы, $a,b\in R^N$, а целевая функция $\varphi(y)$ удовлетворяет the Lipschitz condition
+\begin{equation}\label{Lip}
+\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D.
+\end{equation}
+
+Функция $\varphi(y)$ предполагается многоэкстремальной и заданной в виде ``black box'' (т.е.
+в виде некоторой вычислительной процедуры, на вход которой подается вектор параметров, а на выходе наблюдается соответствующее значение функции).
+Кроме того, подразумевается, что каждое trial (т.е.
+вычисление значение функции в точке области поиска) является трудоемкой операцией и требует значительных вычислительных ресурсов.
+Как отмечалось во введении, такая постановка задачи полностью соответствует the inverse problem of chemical kinetics.
+
+The Lipschitz condition (\ref{Lip}) может быть использовано для оценки глобального минимума функции на интервале, а знание константы Липшица позволяет конструировать алгоритмы глобального поиска и доказывать условия их сходимости (see, for example, [ссылки]).
+
+Одной из основных трудностей в решении многомерных задач глобальной оптимизации является рост вычислительных затрат при увеличении размерности задачи.
+Уменьшение числа испытаний функции при сохранении точности решения возможно за счет более полного использования априорных предположений о целевой функции, что приводит к адаптивным последовательным алгоритмам оптимизации.
+
+Такими методами являются, например, non-uniform space covering method \cite{Evtushenko2013} and simplicial partitions method \cite{Zilinskas2010}.
+Данные подходы были успешно использованы и для построения параллельных оптимизационных методов \cite{Evtushenko2009,Paulavicius2011}.
+Другим адаптивным подходом к решению многомерной задачи (\ref{problemN}) является ее редукция к одной или нескольким одномерным задачам с последующим применением одномерных алгоритмов.
+Такая редукция может быть осуществлена, например, при помощи nested optimization scheme \cite{Grishagin2018} или при помощи Peano-Hilbert curves \cite{Barkalov2018}.
+Последний из указанных подходов и используется в настоящей работе.
+
+Используя непрерывное однозначное отображение (Peano-Hilbert curve) $y(x)$ отрезка $[0,1]$ вещественной оси на гиперинтервал $D$ из (\ref{D}) можно свести многомерную задачу (\ref{problemN}) к одномерной задаче
 \[
-I_k = \{x^{k+1},x^{k+2},...,x^{k+p}\}
+\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}},
 \]
-точек, в которых испытания уже начались, но еще не завершены.
+where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
+\[
+\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}
+\]
+with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation $ H=2 L \sqrt{N+3}$.
+Вопросы численного построения различных аппроксимаций Peano-Hilbert curve рассмотрены в \cite{Strongin2000,Sergeyev2013}.
 
-Step 1. Точки множества 
+Таким образом, поисковое испытание в некотрой точке $x'\in[0,1]$ будет включать в себя сначала the construction of the image $y'=y(x')$, и лишь затем вычисление значения функции $ z' = \varphi(y')$.
+
+\subsection{Parallel Asynchronous Global Search Algorithm}
+
+В предлагаемом подходе схема распараллеливания соответствует по принципу ``мастер-рабочие''.
+В процессе-мастере выполняется алгоритм глобального поиска, в котором проводится накопление поисковой информации, оценка на ее основе константы Липшица для целевой функции, определение точек новых испытаний и распределение их по процессам-рабочим.
+Процессы-рабочие получают от мастера точки, проводят в них новые испытания и отсылают мастеру результаты испытаний.
+
+Будем считать, что на каждой итерации процесс-мастер вычисляет одну точку нового испытания и передает ее для проведения испытания в процесс-рабочий.
+При этом проведение испытания рабочим является значительно более трудоемкой операцией, чем выбор новой точки мастером, что исключает простой процессов-рабочих.
+В данном случае (в отличие от синхронных параллельных алгоритмов) общее число выполненных испытаний каждым процессом-рабочим будет зависеть от трудоемкости проведения конкретного испытания и не может быть оценено заранее.
+
+%комментарий про асинхронность? или ниже?
+При описании параллельного алгоритма будем предполагать, что в нашем распоряжении имеется $p+1$ вычислительный процесс: один мастер и $p$ рабочих.
+
+В начале поиска процесс-мастер (будем считать, что это процесс No 0) инициирует параллельное проведение $p$ испытаний в $p$ различных точках области поиска, две из которых являются граничными, а остальные -- внутренними, т.е.
+в точках $\{y(x^1), y(x^2), ...,y(x^p)\}$, где $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.
+
+Предположим теперь, что выполнено $k$ испытаний (в частности, $k$ может быть равно 0), и процессы-рабочие проводят испытания в точках $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.
+
+Каждый процесс-рабочий, завершивший свое испытание в некоторой точке (не ограничивая общности, будем считать, что это точка $y(x^{k+1})$, соответствующая процессу No 1), пересылает процессу-мастеру результаты испытания.
+Мастер, в свою очередь, выбирает для него точку нового испытания $x^{k+p+1}$ в соответствии с правилами, описанными ниже.
+Отметим, что в данном случае мы будем иметь set of preimages of the trial points
 \[
-X_k
+I_k = \left\{ x^{k+1},x^{k+2},...,x^{k+p} \right\},
 \]
-перенумеровать в порядке возрастания координаты, т.е.
+в которых испытания уже начались, но еще не завершены.
+
+Step 1. Перенумеровать в порядке возрастания (нижним идексом) set of preimages of the trial points 
 \[
-0=
+X_k = \left\{x^1, x^2,...,x^{k+p} \right\},
+\]
+которое содержит все preimages, в которых либо проведены, либо проводятся испытания, т.е.
+\[
+0=x_1<x_2<...<x_{x+p}=1.
 \]
@@
 \]
-где $z_i=\varphi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$. Значения $z_i$ в точках $x_i \in I_k$ являются неопределенными, т.к. испытания в точках $x_i \in I_k$ еще не завершены.
+где $z_i=\varphi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$.
+Значения $z_i$ в точках $x_i \in I_k$ являются неопределенными, т.к.
+испытания в точках $x_i \in I_k$ еще не завершены.
+Если значение $M$ получилось равным 0, то положить $M=1$.
 
-Написать про 0 или неопределеность
+Step 3.
+Каждому интервалу $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, поставить в соответствие число $R(i)$, которое называется характеристикой интервала и вычисляется по формуле
+\[
+R(i)=rM\Delta_i+\frac{(z_i-z_{i-1})^2}{rM\Delta_i}-2(z_i+z_{i-1}),
+\]
+где $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.
 
-Step 3. Каждому интервалу поставить в соответствие число
+Step 4. Выбрать интервал $[x_{t-1},x_t]$, которому соответствует максимальная характеристика, т.е.
+\[
+R(t) = \max \left\{ R(i): \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p \right\}.
+\]
+
+Step 5. Определить точку нового испытания $y^{k+p+1}=y(x^{k+p+1})$, прообраз $x^{k+p+1} \in (x_{t-1},x_t)$ которой в соответствии с формулой
+\[
+x^{k+p+1} = \frac{x_{t}+x_{t-1}}{2} - \mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_{t}-z_{t-1}\right|}{M}\right]^N.
+\]
+
+Сразу после вычисления точки очередного испытания процесс-мастер добавляет ее в множество $I_k$ и пересылает процессу-рабочему, который инициирует проведение испытания в ней.
+
+Процесс-мастер завершает работу алгоритма, если выполняется одно из двух условий: $\Delta_{t}<\epsilon$, $k+p>K_{max}$.
+Вещественное число $\epsilon>0$ и целое число $K_{max}>0$ являются параметрами алгоритма и соответствуют точности поиска решения и максимальному числу испытаний.
+
+Описанный параллельный асинхнонный алгоритм основан на последовательном information global search algorithm.
+Теоретическое обоснование сходимости алгоритма приведено в \cite{Strongin2000}.
+Здесь же приведены схемы синхронного распараллеливания, использовавшиеся ранее при решении ряда прикладных задач \cite{Kalyulin2017,Modorskii2016}.
+Новизна данной работы заключается в практической реализации и применении асинхронной схемы распалаллеливания, обладающей большей эффективностью при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска.
+Это подтверждается результатами экспериментов, описанными в следующем разделе.
 

### 2021-12-08 2022/OM&S/paper.tex
Template for OM&S

% interacttfssample.tex
% v1.05 - August 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

%\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled
%\setlength\bibindent{2em}% To increase hanging indent in bibliography when line spacing is doubled

\usepackage[numbers,sort&compress]{natbib}% Citation support using natbib.sty
\bibpunct[, ]{[}{]}{,}{n}{,}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

\begin{document}

%\articletype{ARTICLE TEMPLATE}% Specify the article type or omit as appropriate

\title{A fast algorithm for finding efficient decisions in multi-objective problems with black-box multiextremal functions}

\author{
\name{Konstantin Barkalov, Vladimir Grishagin and Evgeny Kozinov\thanks{CONTACT Evgeny Kozinov. Email: evgeny.kozinov@itmm.unn.ru}}
\affil{Software Department, Lobachevsky State University, Nizhni Novgorod, Russia}
}

\maketitle

\begin{abstract}
This template is for authors who are preparing a manuscript for a Taylor \& Francis journal using the \LaTeX\ document preparation system and the \texttt{interact} class file, which is available via selected journals' home pages on the Taylor \& Francis website.
\end{abstract}

\begin{keywords}
Multi-objective optimization; global optimization; dimensionality reduction; search information; nature-inspired metaheuristics; numerical comparison
\end{keywords}

\section{Introduction}

\section{Problem statement}

References test \cite{Strongin2000,Grishagin2018}

\section{Algorithm}

\section{Numerical experiments}

\section*{Disclosure statement}

No potential conflict of interest was reported by the authors.

\section*{Funding}

This work was supported by the Russian Science Foundation, project No. 21-11-00204.

\bibliographystyle{tfs}
\bibliography{bibliography}

\end{document}

### 2021-12-08 2022/PAVT/PAVT2022.tex
Template for OM&S

@@
 % Решение обратных задач химической кинетки с помощью смешанного локально-глобального поискового алгоритма 
-\title{Solving the Inverse Problems of Chemical Kinetics Using the Asynchronous Global Optimization Algorithm}
-%
+%\title{Solving the Inverse Problems of Chemical Kinetics Using the Asynchronous Global Optimization Algorithm}
+\title{Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as Catalyst Using the Asynchronous Global Optimization Algorithm}
 \titlerunning{Solving the Inverse Problems of Chemical Kinetics}  % abbreviated title (for running head)
@@
 
-The paper considers the application of parallel computing technology to the simulation of a catalytic chemical reaction, which is widely used in the modern chemical industry to produce synthesis gas.
-As a chemical reaction, the process of pre-reforming propane on a Ni catalyst is assumed.
+The paper considers the application of parallel computing technology to the simulation of catalytic chemical reaction, which is widely used in the modern automobile industry to produce gasoline with high octane number.
+As a chemical reaction, the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid is assumed.
 To simulate a chemical process, it is necessary to develop a kinetic model of the process, that is, to determine the kinetic parameters.
@@
 The use of the parallel algorithm has significantly reduced the search time to find the optimum.
-The found optimal parameters of the model made it possible to adequately simulate the process of pre-reforming propane on a Ni-catalyst.
+The found optimal parameters of the model made it possible to adequately simulate the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid.
 
@@
 %Часть УГНТУ
-
+Currently, there is a tendency to improve the environmental characteristics of automobile fuel while maintaining a high octane number.
+Sulfuric acid alkylation of isobutane with olefins makes it possible to obtain a high-octane component of gasoline with a minimum content of aromatic hydrocarbons.
+The alkylate, which is produced by the alkylation of isobutane with C3-C5 olefins in the presence of strong acid, has the advantages of high octane number, low vapor pressure, and zero content of olefins and aromatics that allow it to be a desirable blending component for high-quality gasoline.
+Alkylates will continue to act as a desirable blending component for high-quality gasoline as the quality of gasoline continues to increase [ссылка на китайскую статью].
+Therefore, it is a significant process of the modern refinery [ссылка на catalyst].
+To optimize the chemical process in industry, it is necessary to first develop its model, which in this case means building a mathematical model of the chemical process and then its kinetic model, that is, numerically calculate the kinetic constants of the reaction.
 %Стыковочный абзац
-Постоянно возрастающая сложность исследуемых реакций влечет за собой усложнение их математических моделей, что существенно затрудняет поиск кинетических констант.
-Как правило, найти кинетические константы реакций аналитическим путем не представляется возможным, поэтому возникает необходимость в разработке и использовании численных методов их поиска (см., например, [ссылки]).
+Как правило, найти кинетические константы реакций аналитическим путем не представляется возможным, поэтому возникает необходимость в разработке и использовании численных методов их поиска (см., например, [ссылки (Наша совместная статья Суперкомп дни 2021 ??)]).
 В этом случае критерий качества найденного решения (целевая функция) не имеет явного аналитического описания, но допускает алгоритмическое представление и требует значительных затрат на расчет.
@@
 \section{Problem Statement}\label{Sec_math_mod}
+%Содержательная постановка задачи
 
-%Содержательная постановка задачи
+Рассмотрим математическую модель реакции алкилирования изобутана олефинами в присутствии серной кислоты, которая представляет собой систему обыкновенных нелинейлых дифференциальных уравнений (\ref{eq:one}) -- (\ref{eq:twelve}).
+
+\begin{equation}
+  \frac{dc_1}{dt} = -k_1c_1 + k_2c_3 - k_3c_1c_3 - k_7c_1c_2c_4 - k_{11}c_1 + k_{14}c_{11}
+  \label{eq:one}
+\end{equation}
+
+\begin{equation}
+  \frac{dc_2}{dt} = -k_4c_2c_4 - k_6c_2c_5 - k_7c_1c_2c_4 - k_{15}c_{11}c_2c_4
+  \label{eq:two}
+\end{equation}
+
+\begin{equation}
+  \frac{dc_3}{dt} = k_1c_1 + k_4c_2c_5 - k_3(c_1 + c_{11})c_3 - k_{5}c_{12}c_3 - k_2c_3 + k_7c_1c_2c_4 + k_{15}c_{11}c_2c_4
+  \label{eq:three}
+\end{equation}
+
+\begin{equation}
+  \frac{dc_4}{dt} = k_3(c_1 + k_{11})c_3 - k_4c_2c_4 - k_{7}c_{1}c_2c_4 - k_{15}c_{11}c_2c_4
+  \label{eq:four}
+\end{equation}
+
+\begin{equation}
+  \frac{dc_5}{dt} = k_5c_{12}c_3 - k_{6}c_2c_5
+  \label{eq:five}
+\end{equation}
+
+\begin{equation}
+  \frac{dc_6}{dt} = k_4c_{2}c_4
+  \label{eq:six}
+\end{equation}
+
+\begin{equation}
+  \frac{dc_7}{dt} = k_6c_{2}c_5 - k_{10}c_7
+  \label{eq:seven}
+\end{equation}
+
+\begin{equation}
+  \frac{dc_8}{dt} = k_7c_{1}c_2c_4 + k_{15}c_{11}c_2c_4 + k_9c_9c_{10} - k_8c_8
+  \label{eq:eight}
+\end{equation}
+
+\begin{equation}
+  \frac{dc_9}{dt} = k_8c_{8} - k_{9}c_{9}c_{10}
+  \label{eq:nine}
+\end{equation}
+
+\begin{equation}
+  \frac{dc_{10}}{dt} = k_8c_{8} - k_{9}c_{9}c_{10}
+  \label{eq:ten}
+\end{equation}
+
+\begin{equation}
+  \frac{dc_{11}}{dt} = -k_3c_{11}c_3 - k_{15}c_{11}c_{2}c_4 + k_{11}c_1 + k_{12}c_{12} - k_{13}c_{11} - k_{14}c_{11}
+  \label{eq:eleven}
+\end{equation}
+
+\begin{equation}
+  \frac{dc_{12}}{dt} = -k_5c_{12}c_3 + k_{13}c_{11} - k_{12}c_{12}
+  \label{eq:twelve}
+\end{equation}
+
+The initial conditions are $t = 0, c_1 = c_1^0; c_2=c_2^0; c_3 = 0; c_4 = 0; c_5= 0; c_6 = 0; c_7 = 0; c_8 = 0; c_9 = 0; c_{10} = 0; c_{11}=c_{11}^0; c_{12} = c_{12}^0$.
+The corresponding species in eqs (\ref{eq:one}) -- (\ref{eq:twelve}) are 1, iC4H8; 2, iC4, 3, iC4+; 4, TMPs+; 5, DMHs+; 6, TMPs; 7, DMHs; 8, HEs; 9, iCx+; 10, iCy=; 11, 2-C4H8; 12, 1-C4H8.
+
+В качестве экспериментальных данных взяты данные со статьи [], которые представляют собой изменение концентраций компонентов реакции во времени при различных температурах; в качестве примера экспериментальных данных для сырья \#1 при температуре 276.2 K представлены данные в таблице \ref{table1}.
+
+\begin{table}
+\label{table1}
+\caption{Experimental Data}
+\begin{center}
+\begin{tabular}{ccccccc}
+\hline
+%\multicolumn{1}{l}{\rule{0pt}{12pt}
+%                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
+& 1 min & 2 min & 5 min & 10 min & 15 min & 20 min  \\
+\hline\rule{0pt}{12pt}
+DMH & 0.12 & 0.11 & 0.1	& 0.1 &	0.095 &	0.09  \\
+TMP & 0.54 & 0.65 & 0.69 & 0.69 & 0.7 & 0.705 \\[2pt]
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+Таким образом, решая систему (\ref{eq:one}) -- (\ref{eq:twelve}) с соответсвующими начальными данными, мы получим изменение рассчитанных концентраций компонентов реакции по времени. ...
+Необходимо минимизировать следующую целевую фукнцию:
+
+\begin{equation}
+  F = \sum ...
+  \label{eq:objective_func}
+\end{equation}
 
@@
 
-%комментарий про асинхронность? или ниже?
 При описании параллельного алгоритма будем предполагать, что в нашем распоряжении имеется $p+1$ вычислительный процесс: один мастер и $p$ рабочих.

### 2021-12-08 2022/PAVT/PAVT2022.tex
before translation of UNN part

@@
 иметь много локальных экстремумов наряду с глобальным.
-
 %Часть ННГУ
@@
 
-\section{}\label{Sec_GSA}
+\section{Parallel Algorithm for Solving Global Optimization Problems }\label{Sec_GSA}
 
@@
 
-%Черновая заготовка для описания параллельного алгоритма оптимизации
 Как уже говорилось выше, с формальной точки зрения the inverse problem of chemical kinetics рассматривается нами как Lipschitz global optimization problem. 
@@
 
-The Lipschitz condition (\ref{Lip}) может быть использовано для оценки глобального минимума функции на интервале, а знание константы Липшица позволяет конструировать алгоритмы глобального поиска и доказывать условия их сходимости (see, for example, [ссылки]).
+The Lipschitz condition (\ref{Lip}) может быть использовано для оценки глобального минимума функции на интервале, а знание константы Липшица позволяет конструировать алгоритмы глобального поиска и доказывать условия их сходимости (see, for example, \cite{Strongin2000}).
 
@@
 
-В предлагаемом подходе схема распараллеливания соответствует по принципу ``мастер-рабочие''.
+В предлагаемом подходе схема распараллеливания соответствует по принципу ``master/worker''.
 В процессе-мастере выполняется алгоритм глобального поиска, в котором проводится накопление поисковой информации, оценка на ее основе константы Липшица для целевой функции, определение точек новых испытаний и распределение их по процессам-рабочим.

### 2021-12-12 2022/PAVT/PAVT2022.tex
UNN part translated

@@
 To optimize the chemical process in industry, it is necessary to first develop its model, which in this case means building a mathematical model of the chemical process and then its kinetic model, that is, numerically calculate the kinetic constants of the reaction.
+
 %Стыковочный абзац
-Как правило, найти кинетические константы реакций аналитическим путем не представляется возможным, поэтому возникает необходимость в разработке и использовании численных методов их поиска (см., например, [ссылки (Наша совместная статья Суперкомп дни 2021 ??)]).
-В этом случае критерий качества найденного решения (целевая функция) не имеет явного аналитического описания, но допускает алгоритмическое представление и требует значительных затрат на расчет.
-Кроме того, в обратных задачах химической кинетики целевая функция может быть существенно многоэкстремальной, т.е.
-иметь много локальных экстремумов наряду с глобальным.
+As a rule, it is impossible to find out the kinetic constants of the reactions analytically.
+Therefore, there is a need in the development and application of numerical methods for finding the kinetic constants (see, e.g., []).
+In this case, the quality criteria of the solution found (objective function) haven't an explicitt analytical description but allows an algorithmic representation and requires considerable computaton resources.
+Moreover, in the inverse problems of chemical kinetics the objective function can be multiextremal essentially, i.e.
+can have many local extrema along with the global one.
+
 %Часть ННГУ
-Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) существенно отличаются от методов локального поиска (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}).
-Как правило, методы локальной оптимизации не могут покинуть области притяжения локального экстремума и не находят глобальный оптимум.
-А использование параметров модели, соответствующих найденному локальному решению, может оказаться недостаточным, поскольку глобальное решение может дать существенный выигрыш по сравнению с локальным.
+The numerical methods for solving such multiextremal problems (global optimization methods) differ significantly from the local search ones (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}).
+As a rule, the local optimization methods cannot escape a local extremum attractione region and don't find the global optimum.
+At the same time, the use of the model parameters corresponding to the local solution found may appear to be insufficient since the global solution may provide a considerable advantage as compared to the local ones.
 
-Разнообразие возникающих задач глобальной оптимизации влечет за собой разнообразие подходов к их решению.
-Методы решения задач глобальной оптмизации можно условно разделить на два класса: метаэвристические и детерминированные.
-Метаэвристические алгоритмы, как правило, основаны на имитации процессов, протекающих в живой природе.
-Примерами метаэвристических алгоритмов являются simulated annealing, evolution and genetic algorithms и т.д.
-(see, e.g., \cite{Battiti2009,Eiben2015}).
-В силу своей относительной простоты метаэвристические алгоритмы более популярны у исследователей, чем детерминированные.
-Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может распологаться далеко от глобального \cite{Kvasov2018}.
+The diversity of the global optimization problems arising entails various approaches to solving these ones.
+The methods of solving the global optimization problems can be divided into two classes: the metaheuristic methods and the deterministic ones.
+The metaheuristic algorithms, as a rule, are based on the simulation of the processes going in the nature.
+Some examples of the metaheuristic algorithms are simulated annealing, evolution and genetic algorithms, etc.
+(see e.g.
+\cite{Battiti2009,Eiben2015}).
+Because of relative simplicity, the metaheuristic algorithms are more popular among the researchers than the deterministic methods.
+However, a problem solution found by a metaheuristic algorithm is, generally speaking, a local one and may be far from the global solution \cite{Kvasov2018}.
 
-Возможность построения детерминированных методов глобального поиска, отличающихся от перебора по сетке и метаэвристических алгоритмов, связана с наличием и учетом  априорных предположений о свойствах функций задачи.
-Такие предположения играют ключевую роль при разработке эффективных алгоритмов глобальной оптимизации и служат основным математическим инструментом для получения оценок глобального решения.
+The possibility to construct deterministic global search methods different from the grid search and from the metaheuristic methods is related to the availability of and taking into account some {\it a priori} assumption on the properties of the problem functions.
+Such assumptions play a key role in the development of efficient global optimization algorithms and serve as main mathematical tools for estimating the global solutions.
 
-Одним из естественных допущений о задаче является предположение об ограниченности относительных изменений значений целевой функции.
-Такое допущение связано с тем, что отношения приращений функции к соответствующим приращениям ее аргумента обычно не могут превышать некоторого порога, определяемого ограниченной энергией изменений в моделируемой системе.
-В таком случае, как известно, функции называются липшицевыми, а сама задача -- задачей липшицевой глобальной оптимизации.
+An assumption on limited relative variations of the objective function values is one of natural assumptions of a problem.
+Such an assumption is related to the ratio of the function increment to respective increment of its argument, which is usually limited by some threshold defined by a limited energy of variations in the simulated system.
+In this case the functions are known as Lipschitz ones and the problem itself is called the Lipschitz global optimization problem.
 
-В данной статье отражены результаты применения параллельных методов липшицевой оптимизации, разрабатываемых в ННГУ им. Н.И. Лобачевского, к решению обратных задач химической кинетики. 
-Основная часть статьи имеет следующую структуру.
-Описание математической модели исследуемой химической реакции приведено в Section 2.
-Формальная постановка задачи липшицевой глобальной оптимизации приводятся в Section 3.
-В Section 4 кратко изложена схема разрабатываемого авторами асинхронного параллельного алгоритма для решения задач указанного класса.
-Результаты численного решения обратной задачи химической кинетики обсуждаются в Section 5.
+This paper presents the results of application of parallel Lipschitz optimization methods for solving the inverse problems of chemical kinetics.
+The main part of the paper has the following structure.
+The description of the mathematical model of the investigated chemical reaction is presented in Section 2.
+The formal statement of the Lipschitz global optimization problems and the asynchronous parallel algorithm for solving these ones are described in Section 3.
+The results of the numerical solving the inverse problem of chemical kinetics are discussed in Section 5.
 
@@
 
-Как уже говорилось выше, с формальной точки зрения the inverse problem of chemical kinetics рассматривается нами как Lipschitz global optimization problem. 
-В общем виде задача указанного класса может быть математически сформулирована следующим образом:
+As it has been already mentioned above, from the formal point of view, we consider the inverse problem of chemical kinetics as Lipschitz global optimization problem.
+In the general form, a problem of the class specified above can be formulated mathematically as
+follows:
 \begin{gather}
@@
 \end{gather}
-где $a,\; b$ есть заданные векторы, $a,b\in R^N$, а целевая функция $\varphi(y)$ удовлетворяет the Lipschitz condition
+where $a,b$ are given vectors, $a,b\in R^N$, and the objective function $\varphi(y)$ satisfies the Lipschitz condition
 \begin{equation}\label{Lip}
@@
 
-Функция $\varphi(y)$ предполагается многоэкстремальной и заданной в виде ``black box'' (т.е.
-в виде некоторой вычислительной процедуры, на вход которой подается вектор параметров, а на выходе наблюдается соответствующее значение функции).
-Кроме того, подразумевается, что каждое trial (т.е.
-вычисление значение функции в точке области поиска) является трудоемкой операцией и требует значительных вычислительных ресурсов.
-Как отмечалось во введении, такая постановка задачи полностью соответствует the inverse problem of chemical kinetics.
+The function $\varphi(y)$ is assumed to be multiextremal and defined in the form of ``black box'' (i.e.
+in the form of some computing procedure, the input of which the  vector of parameters is supplied into, and the corresponding function value is taken from the output).
+Moreover, each \textit{trial} (i.e.
+the computing of the function value at a point of the search domain) is assumed to be time-consuming operation.
+As it has been noted in Introduction, such problem statement corresponds to the inverse problem of chemical kinetics completely.
 
-The Lipschitz condition (\ref{Lip}) может быть использовано для оценки глобального минимума функции на интервале, а знание константы Липшица позволяет конструировать алгоритмы глобального поиска и доказывать условия их сходимости (see, for example, \cite{Strongin2000}).
+The Lipschitz condition (\ref{Lip}) can be utilized to estimate the global minimum of a function within an interval, and knowing the Lipschitz constant allows constructing the global search algorithms and proving the convergence conditions for these ones (see, for example, \cite{Strongin2000}).
 
-Одной из основных трудностей в решении многомерных задач глобальной оптимизации является рост вычислительных затрат при увеличении размерности задачи.
-Уменьшение числа испытаний функции при сохранении точности решения возможно за счет более полного использования априорных предположений о целевой функции, что приводит к адаптивным последовательным алгоритмам оптимизации.
+The growth of the computation costs with increasing problem dimensionality is one of the main difficulties in solving the multidimensional global optimization problems.
+The decreasing of the number of trials at preserving the solution accuracy is possible by a complete utilization of some {\it a priori} assumptions on the objective function that leads to adaptive sequential optimization algorithms.
 
-Такими методами являются, например, non-uniform space covering method \cite{Evtushenko2013} and simplicial partitions method \cite{Zilinskas2010}.
-Данные подходы были успешно использованы и для построения параллельных оптимизационных методов \cite{Evtushenko2009,Paulavicius2011}.
-Другим адаптивным подходом к решению многомерной задачи (\ref{problemN}) является ее редукция к одной или нескольким одномерным задачам с последующим применением одномерных алгоритмов.
-Такая редукция может быть осуществлена, например, при помощи nested optimization scheme \cite{Grishagin2018} или при помощи Peano-Hilbert curves \cite{Barkalov2018}.
-Последний из указанных подходов и используется в настоящей работе.
+For example, non-uniform space covering method \cite{Evtushenko2013} and simplicial partitions method \cite{Zilinskas2010} are such methods.
+These approaches were applied successfully for the development of the parallel optimization methods as well \cite{Evtushenko2009,Paulavicius2011}.
+Another adaptive approach to solving a multidimensional problem (\ref{problemN}) is its reduction to a single one-dimensional problem or to several ones followed by application of the one-dimensional algorithms.
+Such a reduction can be made, for example, using nested optimization scheme \cite{Grishagin2018} or Peano-Hilbert curves \cite{Barkalov2018}. 
+The latter approach was used in the present work.
 
-Используя непрерывное однозначное отображение (Peano-Hilbert curve) $y(x)$ отрезка $[0,1]$ вещественной оси на гиперинтервал $D$ из (\ref{D}) можно свести многомерную задачу (\ref{problemN}) к одномерной задаче
+Using the continuous unambiguous mapping (Peano-Hilbert curve) $y(x)$ of the interval $[0,1]$ of the real axis on the hypercube $D$ from (\ref{D}), one can reduce a multidimensional problem (\ref{problemN}) to a one-dimensional problem
 \[
@@
 with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation $ H=2 L \sqrt{N+3}$. 
-Вопросы численного построения различных аппроксимаций Peano-Hilbert curve рассмотрены в \cite{Strongin2000,Sergeyev2013}.
+The issues of the numerical construction of various approximations of the Peano-Hilbert curve were considered in \cite{Strongin2000,Sergeyev2013}.
 
-Таким образом, поисковое испытание в некотрой точке $x'\in[0,1]$ будет включать в себя сначала the construction of the image $y'=y(x')$, и лишь затем вычисление значения функции $ z' = \varphi(y')$.
+So far, a search trial at some point $x'\in[0,1]$ will include first the construction of the image $y'=y(x')$ and then the computing the value of the function $z'=\varphi(y')$.
 
@@
 
-В предлагаемом подходе схема распараллеливания соответствует по принципу ``master/worker''.
-В процессе-мастере выполняется алгоритм глобального поиска, в котором проводится накопление поисковой информации, оценка на ее основе константы Липшица для целевой функции, определение точек новых испытаний и распределение их по процессам-рабочим.
-Процессы-рабочие получают от мастера точки, проводят в них новые испытания и отсылают мастеру результаты испытаний.
+In the approach proposed, the parallelization scheme corresponds to the ``master/worker'' principle.
+In the master process the global search algorithm is executed, which accumulates the search information, evaluates the Lipschitz constant for the objective function on its base, determines the new trial points and distributes these ones among the worker processes.
 
-Будем считать, что на каждой итерации процесс-мастер вычисляет одну точку нового испытания и передает ее для проведения испытания в процесс-рабочий.
-При этом проведение испытания рабочим является значительно более трудоемкой операцией, чем выбор новой точки мастером, что исключает простой процессов-рабочих.
-В данном случае (в отличие от синхронных параллельных алгоритмов) общее число выполненных испытаний каждым процессом-рабочим будет зависеть от трудоемкости проведения конкретного испытания и не может быть оценено заранее.
+The worker processes receive the trial points from the master process, perform the new trials at these points, and send the trial results to the master process.
 
-При описании параллельного алгоритма будем предполагать, что в нашем распоряжении имеется $p+1$ вычислительный процесс: один мастер и $p$ рабочих.
+Let us assume the master process to compute one point of the next trial at every iteration and to send it to a worker process for executing the trial. 
+At the same time, the execution of the trial by the worker process is much more computational-costly operation than the choice of a new trial point by the master that excludes idle worker processes. 
+In this case (unlike the synchronous parallel algorithms), the total number of trials executed by each worker process will depend on the computation costs of executing a particular trial and cannot be estimated in advance.
 
-В начале поиска процесс-мастер (будем считать, что это процесс No 0) инициирует параллельное проведение $p$ испытаний в $p$ различных точках области поиска, две из которых являются граничными, а остальные -- внутренними, т.е.
-в точках $\{y(x^1), y(x^2), ...,y(x^p)\}$, где $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.
+In the description of the parallel algorithm, let us assume $p+1$ computing processes to be at our disposal: one master process and $p$ worker ones.
 
-Предположим теперь, что выполнено $k$ испытаний (в частности, $k$ может быть равно 0), и процессы-рабочие проводят испытания в точках $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.
+At the beginning of the search, the master process (let us assume it to be Process No 0) initiates the parallel execution of $p$ trials at $p$ different points of the search domain. 
+Two of these points are the boundary ones while the rest are the internal ones, i.e. at the points $\{y(x^1), y(x^2), ...,y(x^p)\}$ where 
+$x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.
 
-Каждый процесс-рабочий, завершивший свое испытание в некоторой точке (не ограничивая общности, будем считать, что это точка $y(x^{k+1})$, соответствующая процессу No 1), пересылает процессу-мастеру результаты испытания.
-Мастер, в свою очередь, выбирает для него точку нового испытания $x^{k+p+1}$ в соответствии с правилами, описанными ниже.
-Отметим, что в данном случае мы будем иметь set of preimages of the trial points
+Now let us assume $k\geq 0$ trials (in particular, $k$ may be equal to 0) to be completed, and the worker processes are performing the trials at the points $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.
+
+Each worker process having completed its trial at some point (without any loss of generality, let us assume this point to be $y(x^{k+1})$ corresponding to process No 1), sends the trial result to the master process.
+In turn, the master process selects a new trial point $x^{k+p+1}$ for the worker process according to the rules described below.
+Note that in this case we will have a set of preimages of the trial points
 \[
@@
 \]
-в которых испытания уже начались, но еще не завершены.
+which the trials have been already started at but haven't been completed yet.
 
-Step 1. Перенумеровать в порядке возрастания (нижним идексом) set of preimages of the trial points 
+Step 1. Renumber set of preimages of the trial points 
 \[
@@
 \]
-которое содержит все preimages, в которых либо проведены, либо проводятся испытания, т.е.
+containing all the preimages, which the trials either have been completed or are being performed at in the increasing order (by the lower index), so that
 \[
@@
 \]
-Step 2. Вычислить значения 
-\[
-M_1=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin I_k, x_i \notin I_k, 2\leq i\leq k+p \right\},
-\]
-\[
-M_2=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in I_k, 2\leq i < k+p \right\},
-\]
-\[
+Step 2. Compute the values 
+\begin{gather*}
+M_1=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin I_k, x_i \notin I_k, 2\leq i\leq k+p \right\},\\
+M_2=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in I_k, 2\leq i < k+p \right\},\\
 M=\max\{M_1,M_2\},
-\]
-где $z_i=\varphi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$.
-Значения $z_i$ в точках $x_i \in I_k$ являются неопределенными, т.к.
-испытания в точках $x_i \in I_k$ еще не завершены.
-Если значение $M$ получилось равным 0, то положить $M=1$.
+\end{gather*}
+where $z_i=\varphi(y(x_i))$ if $x_i \notin I_k, \; 1\leq i \leq k+p$.
+The values $z_i$ at the points $x_i \in I_k$ are undefined since the trials at the points $x_i \in I_k$ haven't been completed yet.
+If the value of $M$ equals 0, then set $M=1$.
 
 Step 3.
-Каждому интервалу $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, поставить в соответствие число $R(i)$, которое называется характеристикой интервала и вычисляется по формуле
+Juxtapose each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$ to a quantity $R(i)$, which is called a characteristic of the interval and is computed according to the formula
 \[
@@
 \]
-где $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.
+where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$ and $r>1$ is the reliability parameter of the method.
 
-Step 4. Выбрать интервал $[x_{t-1},x_t]$, которому соответствует максимальная характеристика, т.е.
+Step 4. Select the interval $[x_{t-1},x_t]$, which the maximum characteristic corresponds to, i.e.
 \[
@@
 
-Step 5. Определить точку нового испытания $y^{k+p+1}=y(x^{k+p+1})$, прообраз $x^{k+p+1} \in (x_{t-1},x_t)$ которой в соответствии с формулой
+Step 5. Define the new trial point $y^{k+p+1}=y(x^{k+p+1})$, the preimage of which is $x^{k+p+1} \in (x_{t-1},x_t)$ according to the formula
 \[
@@
 
-Сразу после вычисления точки очередного испытания процесс-мастер добавляет ее в множество $I_k$ и пересылает процессу-рабочему, который инициирует проведение испытания в ней.
+Upon computing the next trial point, the master process adds it to the set $I_k$ and sends it to the worker process, which initiates the new trial at this point.
 
-Процесс-мастер завершает работу алгоритма, если выполняется одно из двух условий: $\Delta_{t}<\epsilon$, $k+p>K_{max}$.
-Вещественное число $\epsilon>0$ и целое число $K_{max}>0$ являются параметрами алгоритма и соответствуют точности поиска решения и максимальному числу испытаний.
+The master process terminates the algorithm if one of two conditions is satisfied: $\Delta_{t}<\epsilon$ or $k+p>K_{max}$.
+The real number $\epsilon>0$ and the integer number $K_{max}>0$ are the parameters of the algorithm and correspond to the solution search precision and to the maximum number of trials, respectively.
 
-Описанный параллельный асинхнонный алгоритм основан на последовательном information global search algorithm.
-Теоретическое обоснование сходимости алгоритма приведено в \cite{Strongin2000}.
-Здесь же приведены схемы синхронного распараллеливания, использовавшиеся ранее при решении ряда прикладных задач \cite{Kalyulin2017,Modorskii2016}.
-Новизна данной работы заключается в практической реализации и применении асинхронной схемы распалаллеливания, обладающей большей эффективностью при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска.
-Это подтверждается результатами экспериментов, описанными в следующем разделе.
+The parallel asynchronous algorithm described above is based on the sequential information global search algorithm.
+The theoretical substantiation of the algorithm convergence is given in \cite{Strongin2000}.
+Also, the synchronous parallelization schemes used earlier in solving a number of applied problems \cite{Kalyulin2017,Modorskii2016} are presented here.
+The novelty of the present work consists in a practical implementation and application of the asynchronous parallelization scheme featured by a higher efficiency in solving the problems with different computation costs of performing the trials at different points of the search domain.
+It was confirmed by the results of experiments described in the next section.
 

### 2021-12-13 2022/PAVT/PAVT2022.tex
Ufa part in Russian

@@
 %\title{Solving the Inverse Problems of Chemical Kinetics Using the Asynchronous Global Optimization Algorithm}
-\title{Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as Catalyst Using the Asynchronous Global Optimization Algorithm}
-\titlerunning{Solving the Inverse Problems of Chemical Kinetics}  % abbreviated title (for running head)
+\title{Kinetic Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as Catalyst Using the Asynchronous Global Optimization Algorithm}
+\titlerunning{Kinetic Modeling of Isobutane Alkylation}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 
-\begin{equation}
-  \frac{dc_1}{dt} = -k_1c_1 + k_2c_3 - k_3c_1c_3 - k_7c_1c_2c_4 - k_{11}c_1 + k_{14}c_{11}
-  \label{eq:one}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_2}{dt} = -k_4c_2c_4 - k_6c_2c_5 - k_7c_1c_2c_4 - k_{15}c_{11}c_2c_4
-  \label{eq:two}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_3}{dt} = k_1c_1 + k_4c_2c_5 - k_3(c_1 + c_{11})c_3 - k_{5}c_{12}c_3 - k_2c_3 + k_7c_1c_2c_4 + k_{15}c_{11}c_2c_4
-  \label{eq:three}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_4}{dt} = k_3(c_1 + k_{11})c_3 - k_4c_2c_4 - k_{7}c_{1}c_2c_4 - k_{15}c_{11}c_2c_4
-  \label{eq:four}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_5}{dt} = k_5c_{12}c_3 - k_{6}c_2c_5
-  \label{eq:five}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_6}{dt} = k_4c_{2}c_4
-  \label{eq:six}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_7}{dt} = k_6c_{2}c_5 - k_{10}c_7
-  \label{eq:seven}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_8}{dt} = k_7c_{1}c_2c_4 + k_{15}c_{11}c_2c_4 + k_9c_9c_{10} - k_8c_8
-  \label{eq:eight}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_9}{dt} = k_8c_{8} - k_{9}c_{9}c_{10}
-  \label{eq:nine}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_{10}}{dt} = k_8c_{8} - k_{9}c_{9}c_{10}
-  \label{eq:ten}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_{11}}{dt} = -k_3c_{11}c_3 - k_{15}c_{11}c_{2}c_4 + k_{11}c_1 + k_{12}c_{12} - k_{13}c_{11} - k_{14}c_{11}
-  \label{eq:eleven}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_{12}}{dt} = -k_5c_{12}c_3 + k_{13}c_{11} - k_{12}c_{12}
-  \label{eq:twelve}
-\end{equation}
+\begin{gather}
+  \frac{dc_1}{dt} = -k_1c_1 + k_2c_3 - k_3c_1c_3 - k_7c_1c_2c_4 - k_{11}c_1 + k_{14}c_{11} \label{eq:one} \\
+  \frac{dc_2}{dt} = -k_4c_2c_4 - k_6c_2c_5 - k_7c_1c_2c_4 - k_{15}c_{11}c_2c_4 \label{eq:two} \\
+  \frac{dc_3}{dt} = k_1c_1 + k_4c_2c_5 - k_3(c_1 + c_{11})c_3 - k_{5}c_{12}c_3 - k_2c_3 + k_7c_1c_2c_4 + k_{15}c_{11}c_2c_4 \label{eq:three} \\
+  \frac{dc_4}{dt} = k_3(c_1 + k_{11})c_3 - k_4c_2c_4 - k_{7}c_{1}c_2c_4 - k_{15}c_{11}c_2c_4 \label{eq:four} \\
+  \frac{dc_5}{dt} = k_5c_{12}c_3 - k_{6}c_2c_5 \label{eq:five} \\
+  \frac{dc_6}{dt} = k_4c_{2}c_4 \label{eq:six} \\
+  \frac{dc_7}{dt} = k_6c_{2}c_5 - k_{10}c_7 \label{eq:seven} \\
+  \frac{dc_8}{dt} = k_7c_{1}c_2c_4 + k_{15}c_{11}c_2c_4 + k_9c_9c_{10} - k_8c_8 \label{eq:eight} \\
+  \frac{dc_9}{dt} = k_8c_{8} - k_{9}c_{9}c_{10} \label{eq:nine} \\
+  \frac{dc_{10}}{dt} = k_8c_{8} - k_{9}c_{9}c_{10} \label{eq:ten} \\
+  \frac{dc_{11}}{dt} = -k_3c_{11}c_3 - k_{15}c_{11}c_{2}c_4 + k_{11}c_1 + k_{12}c_{12} - k_{13}c_{11} - k_{14}c_{11} \label{eq:eleven} \\
+  \frac{dc_{12}}{dt} = -k_5c_{12}c_3 + k_{13}c_{11} - k_{12}c_{12} \label{eq:twelve} 
+\end{gather}
 
@@
 В качестве экспериментальных данных взяты данные со статьи [], которые представляют собой изменение концентраций компонентов реакции во времени при различных температурах; в качестве примера экспериментальных данных для сырья \#1 при температуре 276.2 K представлены данные в таблице \ref{table1}.
-
 \begin{table}
@@
 \hline
-%\multicolumn{1}{l}{\rule{0pt}{12pt}
-%                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
-& 1 min & 2 min & 5 min & 10 min & 15 min & 20 min  \\
+time, min & 1 & 2 & 5 & 10 & 15 & 20 \\
 \hline\rule{0pt}{12pt}
@@
 
-Таким образом, решая систему (\ref{eq:one}) -- (\ref{eq:twelve}) с соответсвующими начальными данными, мы получим изменение рассчитанных концентраций компонентов реакции по времени. ...
-Необходимо минимизировать следующую целевую фукнцию:
+Таким образом, решая систему (\ref{eq:one}) -- (\ref{eq:twelve}) с соответсвующими начальными данными, мы получим изменение рассчитанных концентраций компонентов реакции по времени.
+Однако, необходимо учитывать тот факт, что константы скоростей реакций $k_1$, $k_2$, ..., $k_{15}$, входящие в уравнения (\ref{eq:one}) -- (\ref{eq:twelve}), являются величинами, зависящими от температуры проведения реакции, причем эта зависимость является уравнением Аррениуса и имеет следующий вид:
 
 \begin{equation}
-  F = \sum ...
-  \label{eq:objective_func}
+  k_i (T) = k_i^0 \exp \left(- \dfrac{E_i}{RT} \right),
+  \label{eq:arren}
 \end{equation}
+where $k_i(T)$ is a constant of $i$-th stage of reation rate, $k_i^0$ is a pre-exponential factor of $i$-th reation stage, $E_i$ is an activation energy, J/mol, $R$ is the universal gas constant, J $\cdot$ (K$\cdot$mol), $T$ is the temperature, K.
+Таким образом, чтобы полностью построить кинетическую модель реакции, необходимо вычислить энергии активации $E_i$ и pre-exponential factors $k_i^0$ всех стадий химической реакции.
+Существует две постановки задач поиска данных кинетических параметров $E$ и $k^0$: первый заключается в решении обратной задачи по подбору кинетических параметров, входящих в уравнение (\ref{eq:arren}), которые в ходе решения позволяют нам вычислить все константы скоростей реакции $k_i$, после чего с найденными константами скоростей стадий реакции проиходит решение системы дифференциальных уравнений (\ref{eq:one}) -- (\ref{eq:twelve}), далее расчитанные концентрации сравниваются с соответсвующими экспериментальными данными.
+Математически, данная задача имеет следующую постановку: необходимо минимизировать следующую целевую фукнцию:
+
+\begin{equation}
+  F_1 = \sum_{i=1}^I \sum_{j=1}^J \sum_{k=1}^K \left| c_{ijk}^{exp} - c_{ijk}^{calc} \right| \longrightarrow \min,
+  \label{eq:objective_func1}
+\end{equation}
+where $c_{ijk}^{exp}$ and $c_{ijk}^{calc}$ are the experimental and calculated values
+of $k$-th observed component at $i$-th experiment, respectively, $I$ is the the number of observed temperatures for the reaction, ($I = 4$ in this case), $K$ is the number of experiments conducted at one temperature ($K = 3$), $M$ is the number of observed components of the reaction ($M = 2$ in this case).
+
+Вторая постановка обратной задачи химической кинетики подразумевает поиск констант скоростей стадии $k_i$, входящих в систему (\ref{eq:one}) -- (\ref{eq:twelve}), отдельно для каждой температуры, далее на основе уравнения Аррениуса (\ref{eq:arren}) с помощью метода наименьших квадратов происходит расчет энергий активации $E_i$ и пред-экспоненциальных множителей $k_i^0$.
+Математически, данная постановка задачи совпадает с уравнением (\ref{eq:objective_func1}) за исключением первого суммирования по температурам:
+
+\begin{equation}
+  F_2 = \sum_{j=1}^J \sum_{k=1}^K \left| c_{ijk}^{exp} - c_{ijk}^{calc} \right| \longrightarrow \min
+  \label{eq:objective_func2} 
+\end{equation}
+
+Таким образом, обе постановки задачи (\ref{eq:objective_func1}) и (\ref{eq:objective_func2}) являются задачами оптимизации и в следующем разделе будет описан метод, который был использован для решения данных задач минимизации.
 
@@
 Juxtapose each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$ to a quantity $R(i)$, which is called a characteristic of the interval and is computed according to the formula
-\[
+\begin{equation}\label{R}
 R(i)=rM\Delta_i+\frac{(z_i-z_{i-1})^2}{rM\Delta_i}-2(z_i+z_{i-1}),
-\]
+\end{equation}
 where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$ and $r>1$ is the reliability parameter of the method.
@@
 
+Согласно постановкам задач (\ref{eq:objective_func1}) и (\ref{eq:objective_func2}) в данной работе были проводились соответсвующие расчеты.
+
+%ННГУ
+Для проведения численных экспериментов использовался UNN supercomputer Lobachevsky (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node).
+Асинхронный алгоритм глобальной оптмизации был реализован на C++ (GCC 5.5.0 and Intel MPI were used); вычисление значений целевой функции было реализовано в Python 3.9.
+
+\subsection{Поиск энергий активации и пред-экспоненциальных множителей реакции}
+
+Сначала решалась задача поиска энергий активации и пред-экспоненциальных множителей всех стадий реакции.
+Количество оптимизируемых параметров равно 30, по два параметра на каждую из пятнадцати стадий реакции.
+Для энергий активации был установлен диапазон поиска $0 \leq E_i \leq 100$ kJ / mol, для пред-экспонециальных множителей -- $0 \leq E_i \leq 10^{12}$ исходя из физико-химических соображений.
+Однако, исходя из жесткости системы дифференциальных уравнений (\ref{eq:one}) и (\ref{eq:twelve}) решение не было найдено в таком широком диапазоне, это связано со слишком высокой степенью пред-экспоненциальных множителей, входящих в уравнение Аррениуса.
+Поэтому в процессе расчетов производилось уменьшение значения верхней границы пред-экспоненциальных множителей и решение было получено при верхней границе $10^5$.
+Найденные кинетические параметры представлены в таблице~\ref{table_res1}.
+
+\begin{table}
+\label{table_res1}
+\caption{Calculated kinetic patameters}
+\begin{center}
+\begin{tabular}{cccccccc}
+\hline
+ & $k_1$ & $k_2$ & $k_3$ & $k_4$ & $k_5$ & $k_6$ & $k_7$\\
+\hline\rule{0pt}{12pt}
+$E$, kJ / mol & 98.60 & 98.22 & 99.07 & 2.29 & 97.61 & 94.30 & 8.67\\
+$k^0$ & $1.46\cdot10^3$ & $1.53\cdot10^2$ & $1.03\cdot10^4$ & $1.10$ & $4.99\cdot10^2$  & $1.31\cdot10^2$ & $5.45\cdot10^2$\\
+\hline
+$k_8$ & $k_9$ & $k_{10}$ & $k_{11}$ & $k_{12}$ & $k_{13}$ & $k_{14}$ & $k_{15}$ \\
+\hline\rule{0pt}{12pt}
+11.57 & 73.16 & 65.54 & 13.86 & 4.93 & 2.69 & 21.19 & 0.63\\
+$2.60\cdot10^2$ & $3.95\cdot10^4$ & $3.24\cdot10^4$ & $1.73\cdot10$ & $3.43$ & $1.78$ & $1.42\cdot10^3$ & $5.57\cdot10$\\[2pt]
+\hline
+\end{tabular}\end{center}\end{table}
+
+%ННГУ
+Для оценки эффективности реализованного метода оптимизации было проведено сравнение результатов работы последовательного, параллельного синхронного и параллельного асинхронного алгоритмов с использованием 8 nodes при решении задачи в данной постановке.
+В таблице \ref{table_30D} приведены минимальные значения objective function, найденные соответствующими методами, время решения задачи (в часах) и ускорение по времени.
+При запуске был использован параметр метода $r=3.0$ из (\ref{R}) и точность $\epsilon = 10^{-2}\left\|b-a\right\|$ в критерии остановки.
+После достижения методом глобального поиска заданной точности решение уточнялось Hook-Jeeves local method \cite{HookJeeves} с точностью $\epsilon = 10^{-4}\left\|b-a\right\|$.
+
+\begin{table}
+\caption{Indicators achieved when solving the problem with 30 parameters}
+\label{table_30D}
+\begin{center}
+\begin{tabular}{cccc}
+\hline\noalign{\smallskip}
+ Method      & Minimum  & Time (h.) & Speedup \\
+\hline\noalign{\smallskip}
+Serial       & 7.6   &    3.8     &  ---        \\
+Synchronous  & 5.9   &   0.9     &   4.3       \\
+Asynchronous & 4.8   &   1.1     &   3.6       \\
+\noalign{\smallskip}\hline
+\end{tabular}\end{center}\end{table}
+
+Результаты показывают, что оба параллельных алгоритма показали невысокое ускорение, но нашли более хорошее решение, чем последовательный метод.
+При этом асинхронный алгоритм нашел лучшее решение, чем синхронный.
+Это также объясняет и меньшее  ускорение асинхронного алгоритма по сравнению с синхронным, т.к.
+асинхронный метод провел больше испытаний в процессе поиска оптимальных кинетических параметров задачи.
+
+%Часть УГНТУ
+С помощью полученных кинетических параметров была решена прямая задача химической кинетики.
+Однако, рассчитанные кинетические кривые плохо описали экспериментальные данные, что подтверждает рисунок \ref{fig:res1}.
+Видно, что характер рассчитанной кривой не совпадает с экспериментальной зависимостью.
+
+\begin{figure}
+\begin{center}
+  \includegraphics[width=0.7\linewidth]{res1.jpg}
+  \caption{Concentration profiles of key components when calculating the activation energies and pre-exponential reaction multipliers.
+Temperature: 276.2 K.
+Symbols, experimental data; line, calculated values.}
+  \label{fig:res1}  
+\end{center}
+\end{figure}
+
+Поэтому было решено провести расчеты согласно второй формулировке задачи (\ref{eq:objective_func2}), а именно вычислить константу каждой стадии реакции по отдельности, затем согласно зависимости Аррениуса найти кинетические параметры $E_i$ и $k_i^0$.
+
+\subsection{Поиск констант скоростей отдельно по каждой температуре}
+
+Далее решалась задача поиска констант $k_i$ скоростей стадии реакции, входящих в систему  (\ref{eq:one}) -- (\ref{eq:twelve}).
+Для каждой из температур были рассчитаны константы (таблица \ref{table_res2}).
+
+\begin{table}
+\label{table_res2}
+\caption{Calculated rate constants}
+\begin{center}
+\begin{tabular}{cccccccc}
+\hline
+T, K & $k_1$ & $k_2$ & $k_3$ & $k_4$ & $k_5$ & $k_6$ & $k_7$\\
+\hline\rule{0pt}{12pt}
+276.2 & 1.66 & 0.13 & 2.09 & 0.08 & 0.11 & 1.83 & 8.82 \\
+279.2 & 1.94 & 0.33 & 2.19 & 0.32 & 0.16 & 1.96 & 14.15\\
+282.2 & 2.14 & 0.99 & 5.43 & 2.77 & 0.56 & 2.13 & 45.85\\
+286.2 & 2.43 & 0.99 & 5.51 & 2.81 & 0.56  & 2.35 & 58.97\\
+\hline
+$k_8$ & $k_9$ & $k_{10}$ & $k_{11}$ & $k_{12}$ & $k_{13}$ & $k_{14}$ & $k_{15}$ \\
+\hline\rule{0pt}{12pt}
+0.87 & 13.38 & 13.71 & 12.57 & 4.20 & 18.54 & 2.70 & 58.61\\
+1.05 & 13.39 & 14.50 & 24.26 & 7.67 & 19.41 & 3.86 & 74.64\\
+1.09 & 18.92 & 17.41 & 33.73 & 6.07 & 19.24 & 9.61 & 77.37\\
+1.50 & 24.09 & 17.88 & 34.22 & 4.35 & 23.04 & 16.38 & 94.45\\[2pt]
+\hline
+\end{tabular}\end{center}\end{table}
+
+%ННГУ
+При решении данной задачи также было проведено сравнение работы последовательного, параллельного синхронного и параллельного асинхронного алгоритмов. 
+В таблице \ref{table_15D} приведены минимальные значения целевой функции, найденные соответствующими методами, время решения задачи (в часах) и ускорение по времени при использовании 8 nodes.
+Все параметры метода были аналогичны предыдущему запуску.
+
+\begin{table}
+\caption{Indicators achieved when solving the problem with 15 parameters}
+\label{table_15D}
+\begin{center}
+\begin{tabular}{cccc}
+\hline\noalign{\smallskip}
+ Method      & Minimum  & Time (h.) & Speedup \\
+\hline\noalign{\smallskip}
+Serial       & 0.35   &   3.4     &  ---        \\
+Synchronous  & 0.36   &   0.4     &   8.6       \\
+Asynchronous & 0.35   &   0.2     &   16.6       \\
+\noalign{\smallskip}\hline
+\end{tabular}\end{center}\end{table}
+
+Результаты показывают, что все алгоритмы находят хорошее по значению целевой функции решение.
+При этом асинхронный алгоритм демонстрирует в два раза большее ускорение, чем синхронный.
+Хорошее ускорение асинхронного алгоритма сохраняется и при большем числе узлов, результаты его работы приведены в таблице \ref{tab_parall}.
+
+\begin{table}
+\caption{Speedup of the asynchronous parallel algorithm}
+\label{tab_parall}
+\begin{center}
+\begin{tabular}{cccc}
+\hline\noalign{\smallskip}
+Nodes  & Minimum  & Time (h.) & Speedup \\
+\hline\noalign{\smallskip}
+1  & 0.35   &   3.4     &   ---        \\
+8  & 0.35   &   0.2     &   16.6       \\
+16 & 0.36   &   0.1     &   34.3       \\
+32 & 0.35   &   0.06    &   59.5       \\
+\noalign{\smallskip}\hline
+\end{tabular}\end{center}\end{table}
+
+%Часть УГНТУ
+C найденными константами решалась прямая задача химической кинетики, результаты сравнения с экспериментальными данными представлены на рисунке \ref{fig:res2} (представлен эксперимент при температуре 276.2 K).
+Видим, что на этот раз описание данных оказалось достаточно точным.
+
+\begin{figure}
+\begin{center}
+  \includegraphics[width=0.7\linewidth]{res2.jpg}
+  \caption{Concentration profiles of key components when calculating the rate constants. Temperature: 276.2 K. Symbols, experimental data; line, calculated values.}
+  \label{fig:res2}  
+\end{center}
+\end{figure}
+
+Рассчитав константы скоростей для различных значений температур, можно вычислить параметры $E_i$ и $k_i^0$.
+Так, например, для первой стадии реакции $E_1 = 24.65$ kJ/mol, $k_1^0 = 6.85 \cdot 10^4$ min$^{-1}$, для третьей стадии $E_3 = 75.57$ kJ/mol, $k_3^0 = 2.62 \cdot 10^{14}$ kg $\cdot$mol$^{-1}\cdot$min$^{-1}$.
+
 \section{Conclusions and Future Work}
+Таким образом, в статье описан поиск кинетических параметров промышленной химической реакции алкилирования изобутана олефинами в присутствии серной кислоты.
+Произведен поиск констант скоростей всех стадий реакции, а также энергий активации и пред-экспоненциальных множителей.
+Метод оптимизации позволил найти досточно точное описание экспериментальных данных.
+В будущем планируется использовать данный метод и распараллеливание для поиска оптимальных условий проведения реакции алкилирования с использованием разработанной кинетической модели.
 

### 2021-12-14 2022/PAVT/PAVT2022.tex
UNN part of Experiments

@@
 Согласно постановкам задач (\ref{eq:objective_func1}) и (\ref{eq:objective_func2}) в данной работе были проводились соответсвующие расчеты.
-
 %ННГУ
-Для проведения численных экспериментов использовался UNN supercomputer Lobachevsky (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node).
-Асинхронный алгоритм глобальной оптмизации был реализован на C++ (GCC 5.5.0 and Intel MPI were used); вычисление значений целевой функции было реализовано в Python 3.9.
+To carry out the numerical experiments, UNN supercomputer Lobachevsky was used (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node).
+The asynchronous global optimization algorithm was implemented using C++ (GCC 5.5.0 and Intel MPI were used); the objective function values were computed using Python 3.9.
 
@@
 %ННГУ
-Для оценки эффективности реализованного метода оптимизации было проведено сравнение результатов работы последовательного, параллельного синхронного и параллельного асинхронного алгоритмов с использованием 8 nodes при решении задачи в данной постановке.
-В таблице \ref{table_30D} приведены минимальные значения objective function, найденные соответствующими методами, время решения задачи (в часах) и ускорение по времени.
-При запуске был использован параметр метода $r=3.0$ из (\ref{R}) и точность $\epsilon = 10^{-2}\left\|b-a\right\|$ в критерии остановки.
-После достижения методом глобального поиска заданной точности решение уточнялось Hook-Jeeves local method \cite{HookJeeves} с точностью $\epsilon = 10^{-4}\left\|b-a\right\|$.
+To evaluate the efficiency of the implemented optimization method, the results obtained using the sequential algorithm, the parallel synchronous, and the parallel asynchronous ones with the use of 8 nodes when solving the problem in the above statement were compared.
+The minimum values of the objective function found by respective methods, the times of solving the problem (in hours), and the values of speedup in time are presented in Table \ref{table_30D}. 
+During the experiments the parameter of the method $r=3.0$ from (\ref{R}) and the accuracy  $\epsilon = 10^{-2}\left\|b-a\right\|$ in the termination condition were used.
+Once the global search method achieves the preset accuracy, the solution was refined by Hook-Jeeves local method \cite{HookJeeves} with the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$.
 
@@
 
-Результаты показывают, что оба параллельных алгоритма показали невысокое ускорение, но нашли более хорошее решение, чем последовательный метод.
-При этом асинхронный алгоритм нашел лучшее решение, чем синхронный.
-Это также объясняет и меньшее  ускорение асинхронного алгоритма по сравнению с синхронным, т.к.
-асинхронный метод провел больше испытаний в процессе поиска оптимальных кинетических параметров задачи.
+The results have shown that both parallel algorithms had demonstrated a moderate speedup but had found better solutions than the sequential method.
+At that, the asynchronous algorithm had found a better solution than synchronous one.
+This explains also a smaller speedup of the asynchronous algorithm as compared with the synchronous one since the asynchronous method had executed more trials in the course of search for the optimal kinetic parameters of the problem.
 
@@
 %ННГУ
-При решении данной задачи также было проведено сравнение работы последовательного, параллельного синхронного и параллельного асинхронного алгоритмов. 
-В таблице \ref{table_15D} приведены минимальные значения целевой функции, найденные соответствующими методами, время решения задачи (в часах) и ускорение по времени при использовании 8 nodes.
-Все параметры метода были аналогичны предыдущему запуску.
+When solving this problem, the operation performances of the sequential algorithm, of the parallel synchronous, and of the parallel asynchronous ones were also compared.  
+The minimum values of the objective function found by respective methods, the time of solving the problem (in hours), and the speedup in time with the use of 8 nodes are presented in Table \ref{table_15D}.
+All the parameters of the method were the same as in the previous run.
 
@@
 
-Результаты показывают, что все алгоритмы находят хорошее по значению целевой функции решение.
-При этом асинхронный алгоритм демонстрирует в два раза большее ускорение, чем синхронный.
-Хорошее ускорение асинхронного алгоритма сохраняется и при большем числе узлов, результаты его работы приведены в таблице \ref{tab_parall}.
+The results have shown that all algorithms had found good solutions (in the values of the objective function).
+At that, the asynchronous algorithm had demonstrated two times higher speedup than the synchronous one.
+A good speedup of the asynchronous algorithm remains at a greater number of nodes as well, the results of its work are presented in Table \ref{tab_parall}.
 

### 2021-12-15 2022/PAVT/PAVT2022.tex
merge UNN and UGNTU parts

@@
 Sulfuric acid alkylation of isobutane with olefins makes it possible to obtain a high-octane component of gasoline with a minimum content of aromatic hydrocarbons.
-The alkylate, which is produced by the alkylation of isobutane with C3-C5 olefins in the presence of strong acid, has the advantages of high octane number, low vapor pressure, and zero content of olefins and aromatics that allow it to be a desirable blending component for high-quality gasoline.
+The alkylate, which is produced by the alkylation of isobutane with C3 -- C5 olefins in the presence of strong acid, has the advantages of high octane number, low vapor pressure, and zero content of olefins and aromatics that allow it to be a desirable blending component for high-quality gasoline.
 Alkylates will continue to act as a desirable blending component for high-quality gasoline as the quality of gasoline continues to increase [ссылка на китайскую статью].
@@
 
-Рассмотрим математическую модель реакции алкилирования изобутана олефинами в присутствии серной кислоты, которая представляет собой систему обыкновенных нелинейлых дифференциальных уравнений (\ref{eq:one}) -- (\ref{eq:twelve}).
+Let us consider a mathematical model of the isobutane alkylation reaction with olefins in the presence of sulfuric acid, which is a system of ordinary nonlinear differential equations (\ref{eq:one}) -- (\ref{eq:twelve}).
 
@@
 
-В качестве экспериментальных данных взяты данные со статьи [], которые представляют собой изменение концентраций компонентов реакции во времени при различных температурах; в качестве примера экспериментальных данных для сырья \#1 при температуре 276.2 K представлены данные в таблице \ref{table1}.
+As experimental data, information is taken from the article [], which represent a change in the concentrations of reaction components over time at different temperatures; as an example of experimental data at a temperature of 276.2 K, the data in the table \ref{table1} are presented.
+
 \begin{table}
@@
 
-Таким образом, решая систему (\ref{eq:one}) -- (\ref{eq:twelve}) с соответсвующими начальными данными, мы получим изменение рассчитанных концентраций компонентов реакции по времени.
-Однако, необходимо учитывать тот факт, что константы скоростей реакций $k_1$, $k_2$, ..., $k_{15}$, входящие в уравнения (\ref{eq:one}) -- (\ref{eq:twelve}), являются величинами, зависящими от температуры проведения реакции, причем эта зависимость является уравнением Аррениуса и имеет следующий вид:
+Thus, solving the system (\ref{eq:one}) -- (\ref{eq:twelve}) with the corresponding initial data, we will get a change in the calculated concentrations of the reaction components over time.
+
+However, it is necessary to take into account the fact that the reaction rate constants $k_1$, $k_2$, ..., $k_{15}$ included in the equations (\ref{eq:one}) -- (\ref{eq:twelve}) are parameters depending on the reaction temperature, and this dependence is the Arrhenius equation and has the following form:
 
@@
 where $k_i(T)$ is a constant of $i$-th stage of reation rate, $k_i^0$ is a pre-exponential factor of $i$-th reation stage, $E_i$ is an activation energy, J/mol, $R$ is the universal gas constant, J $\cdot$ (K$\cdot$mol), $T$ is the temperature, K.
-Таким образом, чтобы полностью построить кинетическую модель реакции, необходимо вычислить энергии активации $E_i$ и pre-exponential factors $k_i^0$ всех стадий химической реакции.
-Существует две постановки задач поиска данных кинетических параметров $E$ и $k^0$: первый заключается в решении обратной задачи по подбору кинетических параметров, входящих в уравнение (\ref{eq:arren}), которые в ходе решения позволяют нам вычислить все константы скоростей реакции $k_i$, после чего с найденными константами скоростей стадий реакции проиходит решение системы дифференциальных уравнений (\ref{eq:one}) -- (\ref{eq:twelve}), далее расчитанные концентрации сравниваются с соответсвующими экспериментальными данными.
-Математически, данная задача имеет следующую постановку: необходимо минимизировать следующую целевую фукнцию:
+Thus, in order to fully develop the kinetic model of the reaction, it is necessary to calculate the activation energies $E_i$ and pre-exponential factors $k_i^0$ of all stages of the chemical reaction.
+There are two formulations of the problems of searching for kinetic parameter data $E$ and $k^0$.
+The first one is to solve the inverse problem of selecting the kinetic parameters included in the equation (\ref{eq:arren}), which during the solution allow us to calculate all the reaction rate constants $k_i$.
+Then, with the found rate constants of the reaction stages, the system of differential equations (\ref{eq:one}) -- (\ref{eq:twelve}) is solved, then the calculated concentrations are compared with the corresponding experimental data.
+Mathematically, this problem has the following formulation: it is necessary to minimize the following objective function:
 
@@
 \end{equation}
-where $c_{ijk}^{exp}$ and $c_{ijk}^{calc}$ are the experimental and calculated values
-of $k$-th observed component at $i$-th experiment, respectively, $I$ is the the number of observed temperatures for the reaction, ($I = 4$ in this case), $K$ is the number of experiments conducted at one temperature ($K = 3$), $M$ is the number of observed components of the reaction ($M = 2$ in this case).
+where $c_{ijk}^{exp}$ and $c_{ijk}^{calc}$ are the experimental and calculated values of $k$-th observed component at $i$-th experiment, respectively, $I$ is the the number of observed temperatures for the reaction, ($I = 4$ in this case), $K$ is the number of experiments conducted at one temperature ($K = 3$), $M$ is the number of observed components of the reaction ($M = 2$ in this case).
 
-Вторая постановка обратной задачи химической кинетики подразумевает поиск констант скоростей стадии $k_i$, входящих в систему (\ref{eq:one}) -- (\ref{eq:twelve}), отдельно для каждой температуры, далее на основе уравнения Аррениуса (\ref{eq:arren}) с помощью метода наименьших квадратов происходит расчет энергий активации $E_i$ и пред-экспоненциальных множителей $k_i^0$.
-Математически, данная постановка задачи совпадает с уравнением (\ref{eq:objective_func1}) за исключением первого суммирования по температурам:
+The second formulation of the inverse problem of chemical kinetics implies the search for the rate constants of the stages $k_i$ included in the system (\ref{eq:one})~--~(\ref{eq:twelve}), separately for each temperature, then on the basis of the Arrhenius equation~(\ref{eq:arren}), the activation energies $E_i$ and the pre-exponential multipliers $k_i^0$ are calculated using the least squares method.
+Mathematically, this statement of the problem coincides with the equation (\ref{eq:objective_func1}), except for the first summation by temperatures:
 
 \begin{equation}
-  F_2 = \sum_{j=1}^J \sum_{k=1}^K \left| c_{ijk}^{exp} - c_{ijk}^{calc} \right| \longrightarrow \min
+  F_2 = \sum_{j=1}^J \sum_{k=1}^K \left| c_{ijk}^{exp} - c_{ijk}^{calc} \right| \longrightarrow \min.
   \label{eq:objective_func2} 
@@
 
-Таким образом, обе постановки задачи (\ref{eq:objective_func1}) и (\ref{eq:objective_func2}) являются задачами оптимизации и в следующем разделе будет описан метод, который был использован для решения данных задач минимизации.
+Thus, both formulations of the problem (\ref{eq:objective_func1}) and (\ref{eq:objective_func2}) are optimization problems and the next section will describe the method that was used to solve these minimization problems.
 
@@
 
-Согласно постановкам задач (\ref{eq:objective_func1}) и (\ref{eq:objective_func2}) в данной работе были проводились соответсвующие расчеты.
+According to the problem statements (\ref{eq:objective_func1}) and (\ref{eq:objective_func2}), the corresponding calculations were carried out in this work.
 %ННГУ
@@
 
-\subsection{Поиск энергий активации и пред-экспоненциальных множителей реакции}
+\subsection{Search for activation energies and pre-exponential multipliers of the reaction}
 
-Сначала решалась задача поиска энергий активации и пред-экспоненциальных множителей всех стадий реакции.
-Количество оптимизируемых параметров равно 30, по два параметра на каждую из пятнадцати стадий реакции.
-Для энергий активации был установлен диапазон поиска $0 \leq E_i \leq 100$ kJ / mol, для пред-экспонециальных множителей -- $0 \leq E_i \leq 10^{12}$ исходя из физико-химических соображений.
-Однако, исходя из жесткости системы дифференциальных уравнений (\ref{eq:one}) и (\ref{eq:twelve}) решение не было найдено в таком широком диапазоне, это связано со слишком высокой степенью пред-экспоненциальных множителей, входящих в уравнение Аррениуса.
-Поэтому в процессе расчетов производилось уменьшение значения верхней границы пред-экспоненциальных множителей и решение было получено при верхней границе $10^5$.
-Найденные кинетические параметры представлены в таблице~\ref{table_res1}.
+First, the problem of searching for activation energies and pre-exponential multipliers of all reaction stages was solved.
+The number of optimized parameters is 30, which is two parameters for each of the fifteen reaction stages.
+For activation energies, the search range was set to $0 \leq E_i \leq 100$ kJ/mol, and $0 \leq E_i \leq 10^{12}$ for pre-exponential multipliers based on physicochemical considerations.
+However, based on the stiffness of the system of differential equations (\ref{eq:one}) and (\ref{eq:twelve}), no solution was found in such a wide range, this is due to the too high degree of pre-exponential multipliers included in the Arrhenius equation.
+Therefore, during the calculations, the value of the upper bound of the pre-exponential multipliers was reduced and the solution was obtained at the upper bounds of $10^5$.
+The kinetic parameters found are presented in the table~\ref{table_res1}.
 
@@
 \label{table_res1}
-\caption{Calculated kinetic patameters}
+\caption{Calculated rate constants of reaction}
 \begin{center}
@@
 %Часть УГНТУ
-С помощью полученных кинетических параметров была решена прямая задача химической кинетики.
-Однако, рассчитанные кинетические кривые плохо описали экспериментальные данные, что подтверждает рисунок \ref{fig:res1}.
-Видно, что характер рассчитанной кривой не совпадает с экспериментальной зависимостью.
+With the help of the obtained kinetic parameters, the direct problem of chemical kinetics was solved.
+However, the calculated kinetic curves poorly described the experimental data, which is confirmed by the figure \ref{fig:res1}.
+It can be seen that the character of the calculated curve does not match the experimental dependence.
 
@@
   \includegraphics[width=0.7\linewidth]{res1.jpg}
-  \caption{Concentration profiles of key components when calculating the activation energies and pre-exponential reaction multipliers.
+  \caption{Concentration profiles of key components when calculating the activation energies and pre-exponential reaction multipliers according to the problem statement~\ref{eq:objective_func1}.
 Temperature: 276.2 K.
@@
 
-Поэтому было решено провести расчеты согласно второй формулировке задачи (\ref{eq:objective_func2}), а именно вычислить константу каждой стадии реакции по отдельности, затем согласно зависимости Аррениуса найти кинетические параметры $E_i$ и $k_i^0$.
+Therefore, it was decided to carry out calculations according to the second formulation of the problem (\ref{eq:objective_func2}), namely, to calculate the constants of each reaction stage separately, then, according to the Arrhenius dependence, find the kinetic parameters $E_i$ and $k_i^0$.
 
-\subsection{Поиск констант скоростей отдельно по каждой температуре}
+\subsection{Searching for rate constants separately for each temperature}
 
-Далее решалась задача поиска констант $k_i$ скоростей стадии реакции, входящих в систему  (\ref{eq:one}) -- (\ref{eq:twelve}).
-Для каждой из температур были рассчитаны константы (таблица \ref{table_res2}).
+Next, the problem of finding constants $k_i$ of the reaction stages included in the system (\ref{eq:one}) -- (\ref{eq:twelve}) was solved.
+Velocity constants were calculated for each of the temperatures (table \ref{table_res2}).
 
@@
 %Часть УГНТУ
-C найденными константами решалась прямая задача химической кинетики, результаты сравнения с экспериментальными данными представлены на рисунке \ref{fig:res2} (представлен эксперимент при температуре 276.2 K).
-Видим, что на этот раз описание данных оказалось достаточно точным.
+The direct problem of chemical kinetics was solved with the found rate constants, the results of comparison with experimental data are shown in Figure \ref{fig:res2} (an experiment at a temperature of 276.2 K is presented).
+We see that this time the description of the data turned out to be rather accurate.
 
@@
   \includegraphics[width=0.7\linewidth]{res2.jpg}
-  \caption{Concentration profiles of key components when calculating the rate constants. Temperature: 276.2 K. Symbols, experimental data; line, calculated values.}
+  \caption{Concentration profiles of key components when calculating the rate constants according to the problem statement~\ref{eq:objective_func2}.
+Temperature: 276.2 K.
+Symbols, experimental data; line, calculated values.}
   \label{fig:res2}  
@@
 
-Рассчитав константы скоростей для различных значений температур, можно вычислить параметры $E_i$ и $k_i^0$.
-Так, например, для первой стадии реакции $E_1 = 24.65$ kJ/mol, $k_1^0 = 6.85 \cdot 10^4$ min$^{-1}$, для третьей стадии $E_3 = 75.57$ kJ/mol, $k_3^0 = 2.62 \cdot 10^{14}$ kg $\cdot$mol$^{-1}\cdot$min$^{-1}$.
+After calculating the rate constants for different temperature values, it is possible to calculate the parameters $E_i$ и $k_i^0$.
+So, for example, for the first stage of the reaction $E_1 = 24.65$ kJ/mol, $k_1^0 = 6.85 \cdot 10^4$ min$^{-1}$, for the third stage $E_3 = 75.57$ kJ/mol, $k_3^0 = 2.62 \cdot 10^{14}$ kg $\cdot$mol$^{-1}\cdot$min$^{-1}$.
+However, for some stages, the parameter values were fairly high $k^0$, namely for the second stage $k_2^0 = 8.52 \cdot 10^{25}$ min$^{-1}$, for the seventh stage $k_7^0 = 2.82 \cdot 10^{26}$ kg mol$^{-2}$ min$^{-2}$.
+Therefore, in future works, the constants corresponding to these kinetic parameters will be recalculated.
 
 \section{Conclusions and Future Work}
-Таким образом, в статье описан поиск кинетических параметров промышленной химической реакции алкилирования изобутана олефинами в присутствии серной кислоты.
-Произведен поиск констант скоростей всех стадий реакции, а также энергий активации и пред-экспоненциальных множителей.
-Метод оптимизации позволил найти досточно точное описание экспериментальных данных.
-В будущем планируется использовать данный метод и распараллеливание для поиска оптимальных условий проведения реакции алкилирования с использованием разработанной кинетической модели.
+Thus, the article describes the search for kinetic parameters of the industrial chemical reaction of isobutane alkylation with olefins in the presence of sulfuric acid.
+A search was made for the rate constants of all reaction stages, as well as activation energies and pre-exponential multipliers have been calculated.
+The optimization method allowed us to find a fairly accurate description of the experimental data.
+In the future, it is planned to use this method and parallelization to eliminate high values of the pre-exponential reaction multipliers and to search for optimal conditions for the alkylation reaction using the developed kinetic model.
 

### 2021-12-15 2022/PAVTOneAPI/OneAPI.tex
Тезисы по OneAPI

\documentclass[11pt, oneside, a4paper]{article}
\usepackage[utf8]{inputenc}
%\usepackage[cp1251]{inputenc} % кодировка
\usepackage[english, russian]{babel} % Русские и английские переносы
\usepackage{graphicx}          % для включения графических изображений
\usepackage{cite}              % для корректного оформления литературы
\usepackage{pavt-ru}  
\usepackage{amsmath}
\usepackage{amssymb}

\begin{document}

% \title - название статьи
% \authors - список авторов

\title{Решение многомерных задач глобальной оптимизации с использованием Intel OneApi\footnote{На что ссылаемся?}}

\authors{К.А. Баркалов\superscript{1}, И.Г. Лебедев\superscript{1} Я.В. Кольтюшкина\superscript{1}}
\organizations{Нижегородский государственный университет им. Н.И. Лобачевского\superscript{1}}

% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

Задача многомерной многоэкстремальной оптимизации может быть определена как поиск наименьшего значения действительной функции \(\phi(y)\)  в гиперинтервале \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\). 
Следует отметить, что в общем случае задача задана по принципу черного ящика. Для целевой функции предполагаем выполнение условия Липшица с априори неизвестной константой \(L\).

В ННГУ им. Н.И. Лобачевского под руководством проф. Р.Г. Стронгина разработан эффективный подход к решению задач глобальной оптимизации \cite{Strongin2013}. В рамках данного подхода решение многомерной задачи сводится к решению одномерной. Для это используется редукция размерности с помощью кривой Пеано \cite{Sergeyev2013}., которая непрерывно и однозначно отображает отрезок вещественной оси \([0,1]\) на  \(N\)-мерный куб.

В процессе работы алгоритма строится последовательность точек \(x^k\), где в каждой точке вычисляется значение минимизируемой функции (производится испытание) \(y^k=f(x^k)\)  Параллельность обеспечиваться следующим образом: в ходе выполнения одной итерации алгоритма будет проводиться \(P\) испытаний одновременно, где \(P \geq 1\).

Одним из вариантов реализации данного подхода является использование инструментов Intel OneAPI. Он прост, открыт и легко интегрируя в существующий код. Intel OneAPI позволяет написать код один раз и в дальнейшем запускать его на различных устройствах.

На данный момент получены результаты вычислительных экспериментов на генераторе тестовых задач GKLS \cite{GKLS}. В таблице \ref{table:GKLS_RES} приведены среднее число итераций параллельного алгоритма глобального поиска и ускорение относительно последовательной версии алгоритма. В каждом эксперименте решалось 100 размерности 4 и 5.

\begin{table}[!hbp]
    \centering
    \caption{Число итераций и ускорение при решении серии 4-х и 5-ти мерных задач}
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    P    & Итераций 4D & ускорение 4D &         Итераций 5D & ускорение 5D \\ \hline
	128 & 106,46   & 4,73      &         235,64   & 4,37      \\ \hline
	256 & 41,25    & 8,50      &         90,45    & 9,20      \\ \hline
	\end{tabular}

    \label{table:GKLS_RES}
\end{table}

% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

\begin{biblio}

\bibitem{Sergeyev2013}
Sergeyev, Y.D., Strongin, R.G., Lera, D. Introduction to global optimization exploiting space-filling curves. 
New York: Springer, 2013. 125 p.

\bibitem{Strongin2013}
Стронгин Р.Г. Гергель В.П. Гришагин В.А. Баркалов К.А. Параллельные вычисления
в задачах глобальной оптимизации.
Москва: Издательство Московского университета,
2013, 280с.

\bibitem{GKLS}
Gaviano M. Lera D. Kvasov D.E. Sergeyev Ya.D. Software for generation of classes of test functions with known local and global minima for global optimization. // ACM Transactions on Mathematical Software. 2003. Vol.~29, No.~4. P.~469–-480.

\end{biblio}
\end{document}

### 2021-12-15 2022/PAVT/PAVT2022.tex
after submission

@@
 Ilya Lebedev$^4$\orcidID{0000-0002-8736-0652} 
+\and
+Dmitry Silenko$^4$ 
 }
@@
 From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem.
-A parallel information-statistical global search algorithm was used to solve it.
-The use of the parallel algorithm has significantly reduced the search time to find the optimum.
+A parallel asynchronous information-statistical global search algorithm was used to solve it.
+The use of the asynchronous algorithm has significantly reduced the search time to find the optimum.
 The found optimal parameters of the model made it possible to adequately simulate the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid.
@@
 The alkylate, which is produced by the alkylation of isobutane with C3 -- C5 olefins in the presence of strong acid, has the advantages of high octane number, low vapor pressure, and zero content of olefins and aromatics that allow it to be a desirable blending component for high-quality gasoline.
-Alkylates will continue to act as a desirable blending component for high-quality gasoline as the quality of gasoline continues to increase [ссылка на китайскую статью].
-Therefore, it is a significant process of the modern refinery [ссылка на catalyst].
+Alkylates will continue to act as a desirable blending component for high-quality gasoline as the quality of gasoline continues to increase \cite{cao2019}.
+Therefore, it is a significant process of the modern refinery.
 To optimize the chemical process in industry, it is necessary to first develop its model, which in this case means building a mathematical model of the chemical process and then its kinetic model, that is, numerically calculate the kinetic constants of the reaction.
@@
 As a rule, it is impossible to find out the kinetic constants of the reactions analytically.
-Therefore, there is a need in the development and application of numerical methods for finding the kinetic constants (see, e.g., []).
+Therefore, there is a need in the development and application of numerical methods for finding the kinetic constants (see, e.g., \cite{Zaynullin2020,Enikeeva2020,Koledina2019,Enikeev2020,uskovfibrous,enikeeva2021,en13133393}).
 In this case, the quality criteria of the solution found (objective function) haven't an explicitt analytical description but allows an algorithmic representation and requires considerable computaton resources.
@@
 The main part of the paper has the following structure.
-The description of the mathematical model of the investigated chemical reaction is presented in Section 2.
-The formal statement of the Lipschitz global optimization problems and the asynchronous parallel algorithm for solving these ones are described in Section 3.
-The results of the numerical solving the inverse problem of chemical kinetics are discussed in Section 5.
+The description of the mathematical model of the investigated chemical reaction is presented in Section \ref{Sec_math_mod}.
+The formal statement of the Lipschitz global optimization problems and the asynchronous parallel algorithm for solving these ones are described in Section \ref{Sec_GSA}.
+The results of the numerical solving the inverse problem of chemical kinetics are discussed in Section \ref{Sec_Exp}.
 
@@
 
-Let us consider a mathematical model of the isobutane alkylation reaction with olefins in the presence of sulfuric acid, which is a system of ordinary nonlinear differential equations (\ref{eq:one}) -- (\ref{eq:twelve}).
+Let us consider a mathematical model of the isobutane alkylation reaction with olefins in the presence of sulfuric acid, which is a system of ordinary nonlinear differential equations (\ref{eq:one})--(\ref{eq:twelve}).
 
@@
 The initial conditions are $t = 0, c_1 = c_1^0; c_2=c_2^0; c_3 = 0; c_4 = 0; c_5= 0; c_6 = 0; c_7 = 0; c_8 = 0; c_9 = 0; c_{10} = 0; c_{11}=c_{11}^0; c_{12} = c_{12}^0$.
-The corresponding species in eqs (\ref{eq:one}) -- (\ref{eq:twelve}) are 1, iC4H8; 2, iC4, 3, iC4+; 4, TMPs+; 5, DMHs+; 6, TMPs; 7, DMHs; 8, HEs; 9, iCx+; 10, iCy=; 11, 2-C4H8; 12, 1-C4H8.
+The corresponding species in eqs (\ref{eq:one})--(\ref{eq:twelve}) are 1, iC4H8; 2, iC4, 3, iC4+; 4, TMPs+; 5, DMHs+; 6, TMPs; 7, DMHs; 8, HEs; 9, iCx+; 10, iCy=; 11, 2-C4H8; 12, 1-C4H8.
 
-As experimental data, information is taken from the article [], which represent a change in the concentrations of reaction components over time at different temperatures; as an example of experimental data at a temperature of 276.2 K, the data in the table \ref{table1} are presented.
+As experimental data, information is taken from the article \cite{cao2019}, which represent a change in the concentrations of reaction components over time at different temperatures; as an example of experimental data at a temperature of 276.2 K, the data in the table \ref{table1} are presented.
 
@@
 
-Thus, solving the system (\ref{eq:one}) -- (\ref{eq:twelve}) with the corresponding initial data, we will get a change in the calculated concentrations of the reaction components over time.
+Thus, solving the system (\ref{eq:one})--(\ref{eq:twelve}) with the corresponding initial data, we will get a change in the calculated concentrations of the reaction components over time.
 
-However, it is necessary to take into account the fact that the reaction rate constants $k_1$, $k_2$, ..., $k_{15}$ included in the equations (\ref{eq:one}) -- (\ref{eq:twelve}) are parameters depending on the reaction temperature, and this dependence is the Arrhenius equation and has the following form:
+However, it is necessary to take into account the fact that the reaction rate constants $k_1, k_2, ..., k_{15}$ included in the equations (\ref{eq:one})--(\ref{eq:twelve}) are parameters depending on the reaction temperature, and this dependence is the Arrhenius equation and has the following form:
 
@@
 The first one is to solve the inverse problem of selecting the kinetic parameters included in the equation (\ref{eq:arren}), which during the solution allow us to calculate all the reaction rate constants $k_i$.
-Then, with the found rate constants of the reaction stages, the system of differential equations (\ref{eq:one}) -- (\ref{eq:twelve}) is solved, then the calculated concentrations are compared with the corresponding experimental data.
+Then, with the found rate constants of the reaction stages, the system of differential equations (\ref{eq:one})--(\ref{eq:twelve}) is solved, then the calculated concentrations are compared with the corresponding experimental data.
 Mathematically, this problem has the following formulation: it is necessary to minimize the following objective function:
@@
 \end{equation}
-where $c_{ijk}^{exp}$ and $c_{ijk}^{calc}$ are the experimental and calculated values of $k$-th observed component at $i$-th experiment, respectively, $I$ is the the number of observed temperatures for the reaction, ($I = 4$ in this case), $K$ is the number of experiments conducted at one temperature ($K = 3$), $M$ is the number of observed components of the reaction ($M = 2$ in this case).
+where $c_{ijk}^{exp}$ and $c_{ijk}^{calc}$ are the experimental and calculated values of $k$-th observed component at $i$-th experiment, respectively, $I$ is the the number of observed temperatures for the reaction, ($I = 4$ in this case), $K$ is the number of experiments conducted at one temperature ($K = 3$), $M$ is the number of observed components of the reaction ($M = 2$).
 
-The second formulation of the inverse problem of chemical kinetics implies the search for the rate constants of the stages $k_i$ included in the system (\ref{eq:one})~--~(\ref{eq:twelve}), separately for each temperature, then on the basis of the Arrhenius equation~(\ref{eq:arren}), the activation energies $E_i$ and the pre-exponential multipliers $k_i^0$ are calculated using the least squares method.
+The second formulation of the inverse problem of chemical kinetics implies the search for the rate constants of the stages $k_i$ included in the system (\ref{eq:one})--(\ref{eq:twelve}), separately for each temperature, then on the basis of the Arrhenius equation~(\ref{eq:arren}), the activation energies $E_i$ and the pre-exponential multipliers $k_i^0$ are calculated using the least squares method.
 Mathematically, this statement of the problem coincides with the equation (\ref{eq:objective_func1}), except for the first summation by temperatures:
@@
 The minimum values of the objective function found by respective methods, the times of solving the problem (in hours), and the values of speedup in time are presented in Table \ref{table_30D}. 
-During the experiments the parameter of the method $r=3.0$ from (\ref{R}) and the accuracy  $\epsilon = 10^{-2}\left\|b-a\right\|$ in the termination condition were used.
-Once the global search method achieves the preset accuracy, the solution was refined by Hook-Jeeves local method \cite{HookJeeves} with the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$.
+During the experiments the parameter of the method $r=3.0$ from (\ref{R}) and the accuracy  $\epsilon = 10^{-3}\left\|b-a\right\|$ in the termination condition were used.
+Once the global search method achieves the preset accuracy, the solution was refined by Hook-Jeeves local method \cite{HookJeeves} with the accuracy $\epsilon = 10^{-5}\left\|b-a\right\|$.
 
@@
 With the help of the obtained kinetic parameters, the direct problem of chemical kinetics was solved.
-However, the calculated kinetic curves poorly described the experimental data, which is confirmed by the figure \ref{fig:res1}.
+However, the calculated kinetic curves poorly described the experimental data, which is confirmed by Fig.
+\ref{fig:res1}.
 It can be seen that the character of the calculated curve does not match the experimental dependence.
@@
 
-Next, the problem of finding constants $k_i$ of the reaction stages included in the system (\ref{eq:one}) -- (\ref{eq:twelve}) was solved.
-Velocity constants were calculated for each of the temperatures (table \ref{table_res2}).
+Next, the problem of finding constants $k_i$ of the reaction stages included in the system (\ref{eq:one})--(\ref{eq:twelve}) was solved.
+Velocity constants were calculated for each of the temperatures (Table \ref{table_res2}).
 
-\begin{table}
-\label{table_res2}
-\caption{Calculated rate constants}
+\begin{table}[ht]
+\caption{Calculated rate constants}\label{table_res2}
 \begin{center}
@@
 %Часть УГНТУ
-The direct problem of chemical kinetics was solved with the found rate constants, the results of comparison with experimental data are shown in Figure \ref{fig:res2} (an experiment at a temperature of 276.2 K is presented).
+The direct problem of chemical kinetics was solved with the found rate constants, the results of comparison with experimental data are shown in Fig.
+\ref{fig:res2} (an experiment at a temperature of 276.2 K is presented).
 We see that this time the description of the data turned out to be rather accurate.
@@
 
-After calculating the rate constants for different temperature values, it is possible to calculate the parameters $E_i$ и $k_i^0$.
+After calculating the rate constants for different temperature values, it is possible to calculate the parameters $E_i$ and $k_i^0$.
 So, for example, for the first stage of the reaction $E_1 = 24.65$ kJ/mol, $k_1^0 = 6.85 \cdot 10^4$ min$^{-1}$, for the third stage $E_3 = 75.57$ kJ/mol, $k_3^0 = 2.62 \cdot 10^{14}$ kg $\cdot$mol$^{-1}\cdot$min$^{-1}$.
@@
 
-\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204 and by RFBR, project No.\,19-37-60014.
+\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204, and by RFBR, project No.\,19-37-60014.
 

### 2021-12-15 2022/PAVTOneAPI/OneAPI.tex
Поправлена ссылка на грант

@@
 
-\title{Решение многомерных задач глобальной оптимизации с использованием Intel OneApi\footnote{На что ссылаемся?}}
+\title{Решение многомерных задач глобальной оптимизации с использованием Intel OneApi\footnote{Исследование выполнено при финансовой поддержке РФФИ в рамках научного проекта 19-07-00242.}}
 

### 2021-12-15 2022/PAVTOneAPI/OneAPI.tex
PAVT poster

--- 2021-12-15 2022/PAVTOneAPI/OneAPI.tex
@@
 
-\title{Решение многомерных задач глобальной оптимизации с использованием Intel OneApi\footnote{На что ссылаемся?}}
+\title{Решение многомерных задач глобальной оптимизации с использованием Intel OneApi\footnote{Исследование выполнено за счет гранта РНФ № 21-11-00204.}}
 
@@
 
-Задача многомерной многоэкстремальной оптимизации может быть определена как поиск наименьшего значения действительной функции \(\phi(y)\)  в гиперинтервале \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
+Задача многомерной многоэкстремальной оптимизации может быть определена как поиск наименьшего значения действительной функции \(\varphi(y)\)  в гиперинтервале \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
 Следует отметить, что в общем случае задача задана по принципу черного ящика. Для целевой функции предполагаем выполнение условия Липшица с априори неизвестной константой \(L\).
@@
 Н.И.
-Лобачевского под руководством проф.
-Р.Г.
-Стронгина разработан эффективный подход к решению задач глобальной оптимизации \cite{Strongin2013}.
+Лобачевского разработан эффективный подход к решению задач глобальной оптимизации \cite{Strongin2013}.
 В рамках данного подхода решение многомерной задачи сводится к решению одномерной.
-Для это используется редукция размерности с помощью кривой Пеано \cite{Sergeyev2013}., которая непрерывно и однозначно отображает отрезок вещественной оси \([0,1]\) на  \(N\)-мерный куб.
+Для это используется редукция размерности с помощью кривой Пеано $y(x)$, которая непрерывно и однозначно отображает отрезок вещественной оси \([0,1]\) на  \(N\)-мерный куб \cite{Sergeyev2013}.
 
-В процессе работы алгоритма строится последовательность точек \(x^k\), где в каждой точке вычисляется значение минимизируемой функции (производится испытание) \(y^k=f(x^k)\)  Параллельность обеспечиваться следующим образом: в ходе выполнения одной итерации алгоритма будет проводиться \(P\) испытаний одновременно, где \(P \geq 1\).
+В процессе работы алгоритма строится последовательность точек \(x^k\), в которых поводятся \textit{испытания}, т.е.
+вычисляются значения целевой функции \(\varphi^k=\varphi(y(x^k))\).
+Распараллеливание организовано следующим образом: в ходе выполнения одной итерации алгоритма одновременно проводится \(P \geq 1\) испытаний в различных точках области поиска.
 
@@
 В таблице \ref{table:GKLS_RES} приведены среднее число итераций параллельного алгоритма глобального поиска и ускорение относительно последовательной версии алгоритма.
-В каждом эксперименте решалось 100 размерности 4 и 5.
+В каждом эксперименте решалось 100 размерности $N=4$ и $N=5$.
 
@@
     \hline
-    P    & Итераций 4D & ускорение 4D &         Итераций 5D & ускорение 5D \\ \hline
-	128 & 106,46   & 4,73      &         235,64   & 4,37      \\ \hline
-	256 & 41,25    & 8,50      &         90,45    & 9,20      \\ \hline
+    P    & Итерации ($N=4$) & Ускорение ($N=4$) &         Итерации ($N=5$) & Ускорение ($N=5$) \\ \hline
+	128 & 106.5   & 4.7      &         235.6   & 4.4      \\ \hline
+	256 & 41.2    & 8.5      &         90.4    & 9.2      \\ \hline
 	\end{tabular}
@@
 
-\bibitem{Sergeyev2013}
-Sergeyev, Y.D., Strongin, R.G., Lera, D. Introduction to global optimization exploiting space-filling curves. 
-New York: Springer, 2013. 125 p.
-
 \bibitem{Strongin2013}
@@
 Москва: Издательство Московского университета,
-2013, 280с.
+2013, 280 с.
+
+\bibitem{Sergeyev2013}
+Sergeyev Y.D., Strongin R.G., Lera D. Introduction to global optimization exploiting space-filling curves. 
+New York: Springer, 2013, 125 p.
 
 \bibitem{GKLS}
-Gaviano M.
-Lera D.
-Kvasov D.E.
-Sergeyev Ya.D.
-Software for generation of classes of test functions with known local and global minima for global optimization.
-// ACM Transactions on Mathematical Software.
+Gaviano M., Lera D., Kvasov D.E., Sergeyev Ya.D.
+Software for generation of classes of test functions with known local and global minima for global optimization // ACM Transactions on Mathematical Software.
 2003.
-Vol.~29, No.~4.
-P.~469–-480.
+vol.~29(4).
+p.~469--480.
 

### 2022-01-02 2022/OM&S/paper.tex
Introduction

@@
 \documentclass[]{interact}
+
+\usepackage{color}
+\usepackage{hyperref}
 
@@
 
+In the field of mathematical modeling the intelligent decision-making processes, the problems of multicriterial optimization (MCO) present a wide class of models that are rich and interesting from a theoretical point of view and very important in real applications.
+Multicriterial problems describe decision-making situations with multiple goals which are considered as functional criteria to be optimized.
+
+In general case, there exist three key factors determining the complexity of analyzing these models.
+The crucial feature corresponds to the typical case where criteria are contradictory: increasing one criterion leads to decreasing the other.
+Such inconsistent behavior generates the necessity of introducing a compromise in the joint analysis of criteria and consideration of a set of compromised (partial) solutions as a complete solution of a multiobjective problem (Pareto set).
+The second complexity factor is associated with the dimensionality or number of varied parameters of the problem because increasing the problem dimension leads to significant growth of computational cost.
+
+Finally, the last factor is the multiextremality of criteria.
+First of all, multiextremal criteria can generate complicated Pareto sets consisting, for example, of several disconnected parts.
+Moreover, the combination of such factors as multiextremality and greater dimension leads to significant computational complexity of the problem because the computational cost increases exponentially when dimension grows.
+From this point of view the multiextremal problems are the most complicated ones in multicriterial optimization and this MCO class is the subject of interest in the given paper.
+
+The simplest way to build an approximation of the Pareto set consist in immersion in the feasible domain of problem's parameters a uniform grid (deterministic or random) and selection of all non-dominated nodes of the grid.
+Unfortunately, this universal approach is too costly because it requires in the worst case pairwise comparison of all the grid nodes and does not take into account the properties of the problem for reducing the computational costs.
+
+Another idea applies a parameterization scheme for the criteria of the problem and reduces a MCO problem to a family of single-criterion (scalar) optimization problems being problems of mathematical programming which can be solved by corresponding well-developed optimization methods.
+In this scheme a non-negative coefficient (weight) is assigned to each criterion which reflects numerically an ``importance'' of the criterion and on the base of such weighted criteria a scalar function (convolution) is built and minimized.
+Under appropriate choice of convolution the solution of the built scalar optimization problem is a partial solution of the initial multicriterial problem.
+Changing weights we can obtain different Pareto points.
+
+There exist various convolution functions but weighted sum function (linear convolution) and maximum from weighted criteria (maximum convolution) are most commonly used.
+The latter is more universal because it enables to build completely the Pareto set taking all the possible combinations of weights, whereas the linear convolution can lose Pareto points in non-convex case (in particular, if criteria are multiextremal).
+
+For optimization of convolutions a wide spectrum of algorithms can be used in dependence on convolution properties generated by the features of criteria in the initial MCO problem.
+In particular, for the multiextremal case where it is required to find the global optimum the publications \cite{Evtushenko2014,Zilinskas2015,GERGEL2017_1,Gergel2019_2,Barkalov2021} describe applications of the global optimization algorithms to MCO problems in combination with the convolution approach.
+
+Among other methods of solving MCO problems it is necessary to pay attention to the important class of metaheuristic nature-inspired algorithms.
+Many of them are nature-inspired, i.e.,  are  based on modeling physical processes (for example, simulated annealing \cite{Locatelli2002,Aarts2014}) or mainly behavior of biological agents including evolutionary \cite{Price2005,Coello2007} and as their part genetic algorithms \cite{Ruiz2015,Deb2002,Zitzler2001} and particle swarm methods modeling the swarming intelligence of bees, fireflies, birds, ants, etc.
+\cite{Mostaghim2007,Nebro2009,Durillo2010}.
+
+Consideration of different approaches to solving MCO problems and a relevant references can be found, for example, in monographs and papers \cite{Zhou2011,Nedjah2015} \colorbox[rgb]{1,1,0}{[1-  Zhou, Nedjah ]}.
+In this paper we consider multiextremal black-box MCO problems and propose a new scheme of solving these problems.
+This scheme is based on maximum convolution, dimensionality reduction by means of Peano curves, information-statistical global optimization algorithm used for the first time in MCO problems.
+Moreover, in the framework of the scheme the information already obtained in the course of solving scalar subproblems of convolution optimization can be utilized for solving a new convolution subproblem.
+Briefly, the way is as follows.
+Before beginning the optimization of a new convolution we have at our disposal values of criteria computed already and now can recalculate them to values of the current convolution taking into account this values as initial information for optimization algorithm.
+This procedure accelerates significantly the optimization process.
+
+The quality of the proposed MCO method is experimentally studied on several multidimensional multiextremal test problems.
+Comparison with known evolutionary algorithms (2 particle swarm methods and 3 genetic algorithms have been taken) on the base of the hypervolume index \cite{Evtushenko2014,Zilinskas2015} reflecting quality of Pareto set estimation demonstrates qualitative results of the proposed method.
+
+The rest of the paper is organized in the following manner.
+Section 2 considers the statement of the MCO problem to be studied.
+Section 3 is devoted to the description of the global search algorithm applied for scalar optimization of convolutions and the scheme of joint analyzing the family of ones.
+Section 4 contains results of computational experiments.
+The last Section concludes the paper.
+
 \section{Problem statement}

### 2022-01-02 2022/OM&S/paper.tex
include more references in the introduction

@@
 
-\usepackage{color}
+%\usepackage{color}
 \usepackage{hyperref}
@@
 
-In the field of mathematical modeling the intelligent decision-making processes, the problems of multicriterial optimization (MCO) present a wide class of models that are rich and interesting from a theoretical point of view and very important in real applications.
+In the field of mathematical modeling the intelligent decision-making processes, the problems of multicriterial optimization (MCO) present a wide class of models that are rich and interesting from a theoretical point of view and very important in real applications \cite{Marler2009,Hillermeier2005}.
 Multicriterial problems describe decision-making situations with multiple goals which are considered as functional criteria to be optimized.
@@
 The crucial feature corresponds to the typical case where criteria are contradictory: increasing one criterion leads to decreasing the other.
-Such inconsistent behavior generates the necessity of introducing a compromise in the joint analysis of criteria and consideration of a set of compromised (partial) solutions as a complete solution of a multiobjective problem (Pareto set).
+Such inconsistent behavior generates the necessity of introducing a compromise in the joint analysis of criteria and consideration of a set of compromised (partial) solutions as a complete solution of a multiobjective problem (Pareto set), see, e.g., \cite{Collette2004,Ehrgott2005}.
 The second complexity factor is associated with the dimensionality or number of varied parameters of the problem because increasing the problem dimension leads to significant growth of computational cost.
 
-Finally, the last factor is the multiextremality of criteria.
+Finally, the last factor is the multiextremality of criteria \cite{Pardalos2017}.
 First of all, multiextremal criteria can generate complicated Pareto sets consisting, for example, of several disconnected parts.
@@
 
-Another idea applies a parameterization scheme for the criteria of the problem and reduces a MCO problem to a family of single-criterion (scalar) optimization problems being problems of mathematical programming which can be solved by corresponding well-developed optimization methods.
+Another idea applies a parameterization scheme for the criteria of the problem and reduces a MCO problem to a family of single-criterion (scalar) optimization problems being problems of mathematical programming which can be solved by corresponding well-developed optimization methods \cite{Collette2004,Ehrgott2005}.
 In this scheme a non-negative coefficient (weight) is assigned to each criterion which reflects numerically an ``importance'' of the criterion and on the base of such weighted criteria a scalar function (convolution) is built and minimized.
@@
 
-Consideration of different approaches to solving MCO problems and a relevant references can be found, for example, in monographs and papers \cite{Zhou2011,Nedjah2015} \colorbox[rgb]{1,1,0}{[1-  Zhou, Nedjah ]}.
+Consideration of different approaches to solving MCO problems and a relevant references can be found, for example, in monographs and papers \cite{Miettinen1999,Ehrgott2005,Zhou2011,Nedjah2015,Pardalos2017}.
 In this paper we consider multiextremal black-box MCO problems and propose a new scheme of solving these problems.

### 2022-01-04 2022/OM&S/paper.tex
Section 2

@@
 Briefly, the way is as follows.
-Before beginning the optimization of a new convolution we have at our disposal values of criteria computed already and now can recalculate them to values of the current convolution taking into account this values as initial information for optimization algorithm.
+Before beginning the optimization of a new convolution we have at our disposal values of criteria computed already and now can recalculate them to values of the current convolution taking into account these values as initial information for optimization algorithm.
 This procedure accelerates significantly the optimization process.
@@
 
-\section{Problem statement}
+\section{MCO problem statement}
 
-References test \cite{Strongin2000,Grishagin2018}
+The model of decision making to be considered hereinafter as the MCO problem consists in the following.
+
+Let $w_i(x), 1 \leq i \leq k,$ be real-valued functions depending on vector of arguments $x=(x_1,...,x_N)$  and defined over the hyperparallelepiped 
+\begin{equation}\label{Q}
+Q = \left\{x \in R^N : a_j \leq x_j \leq b_j, 1 \leq i \leq N \right\}
+\end{equation}
+in $N$-dimensional Euclidean space $R^N$.
+These functions reflect objectives of decision making and it is considered that the less the function value, the better the result of achieving this objective.
+
+The problem of multicriterial optimization is formulated as minimizing over the domain (\ref{Q}) the vector function
+\begin{equation}\label{W}
+W(x) = \left(w_1(x), ..., w_k(x)\right).
+\end{equation}
+
+This problem can be symbolically written as
+\begin{equation}\label{problem}
+W(x) \rightarrow \min , \; x \in Q.
+\end{equation}
+
+In terminology of decision making the vector function $W(x)$  is called \textit{vector criterion} of the problem, functions $w_i(x), 1 \leq i \leq k,$ are \textit{partial criteria} or \textit{objective functions}, $Q$ is the domain of feasible decisions or just the \textit{feasible domain} and points $x \in Q$ are \textit{feasible decisions}.
+
+As a rule, each criterion attains its minimum at a point different from minimum points of other criteria, i.e., it is impossible to find out the decision $x^*$   providing in the region $Q$ minimum values for all the partial criteria simultaneously.
+This situation generates the contradictoriness of criteria, when decreasing one criterion leads to the growth of the other, and complicates the notion of optimal decision (solution to the problem (\ref{problem})).
+
+Traditionally, optimal decisions in the problem (\ref{problem}) are defined on the base of Pareto optimality concept.
+To complete this item, let us give the known classical definitions concerning this concept.
+
+\begin{definition} 
+Let $x,z \in Q$.
+Vector $x$ is said to \textit{dominate} vector $z$ ($x \succ z$) if $w_i(x) \leq w_i(z), \; 1 \leq i \leq k,$ and there exists a number $p, \; 1 \leq p \leq k,$  such that $w_p(x) < w_p(z)$.
+\end{definition}
+\begin{definition} 
+A decision vector $x^* \in Q$ is a \textit{Pareto optimal point} if in the domain $Q$ there is no vector $z \neq x^*$ dominating $x^*$.
+\end{definition}
+\begin{definition} 
+The set of all Pareto optimal points (Pareto set $P$) is the \textit{Pareto optimal solution} to the MCO problem (\ref{problem}).
+\end{definition}
+\begin{definition} 
+The set $F=W(P)=\left\{W(x):x \in P\right\}$  is called the \textit{Pareto front} of the problem (\ref{problem}).
+\end{definition}
+
+As mentioned in Introduction, one way of building the Pareto set consists in reducing the multiobjective problem to a family of single-criteria, or scalar optimization problems and solving them by methods of mathematical programming.
+The global minimum point of a problem from this family will be a Pareto optimal point under corresponding assumptions.
+This approach can be realized by means of applying a convolution of the vector criterion (\ref{W}), for example, the maximum convolution
+\begin{equation}\label{conv}
+\Gamma(\lambda,x) = \max_{1 \leq i \leq k}{\lambda_i w_i(x)},
+\end{equation}
+where coefficients $\lambda_i, \; 1 \leq i \leq k,$ (weights of criteria) satisfy the conditions
+\begin{equation}\label{lambda}
+\lambda_i \geq 0, \; 1 \leq i \leq k, \; \sum_{i=1}^k{\lambda_i} = 1.
+\end{equation}
+
+If
+\begin{equation}\label{positive}
+w_i(x) > 0, \; 1 \leq i \leq k,
+\end{equation}
+in the domain (\ref{Q}), then global solution to the problem
+\begin{equation}\label{conv_problem}
+\Gamma(\lambda,x) + \gamma \sum_{i=1}^k{\lambda_i w_i(x)} \rightarrow \min, \; x \in Q,
+\end{equation}
+with a sufficiently small parameter $\gamma > 0$ is a Pareto optimal point of the initial MCO problem (\ref{problem}) \cite{Wierzbicki} [Wierzbicki, Krasnoshekov].
+The positiveness of criteria is not restrictive requirement because it is possible to transform easily the MCO problem with non-positive criteria to the form (\ref{positive}) without loss of Pareto solution.
+
+In the case of multiextremal criteria the problem (\ref{conv_problem}) is multiextremal as well and it is necessary to use global optimization algorithms for its solving.
+Many references to such the algorithms can be found in monographs \cite{Strongin2000,Pinter1996,Zhigljavsky2008,Sergeyev2013,PaulaviciusZilinskas2014,Sergeyev2017}.
+For solving multiextremal problems the information-statistical algorithms \cite{Strongin2000,Sergeyev2013} are among of the most efficient.
+
+In the next section we describe an information-statistical algorithm with accelerated convergence and its application to searching for Pareto optimal solutions in multiextremal MCO problems.
 

### 2022-01-06 2022/OM&S/paper.tex
Section 3

@@
 \begin{equation}\label{conv_problem}
-\Gamma(\lambda,x) + \gamma \sum_{i=1}^k{\lambda_i w_i(x)} \rightarrow \min, \; x \in Q,
+\Phi_\lambda(x) = \Gamma(\lambda,x) + \gamma \sum_{i=1}^k{\lambda_i w_i(x)} \rightarrow \min, \; x \in Q,
 \end{equation}
@@
 
-\section{Algorithm}
+\section{Computational scheme for multiextremal multiobjective optimization}
+
+In this section we consider the MCO problem under additional assumption according to which the criteria $w_i(x), \; 1 \leq i \leq k,$ are black-box functions and satisfy in the domain $Q$ the Lipschitz condition
+\begin{equation}\label{lip}
+\left| w_i(x) - w_i(y)\right| \leq L_i \left\| x-y \right\|, \; x,y, \in Q, \; 1 \leq i \leq k,
+\end{equation}
+with corresponding Lipchitz constants $L_i > 0$ that are, as a rule, unknown. Here $\left\|*\right\|$   denotes the Euclidean norm in $R^N$.
+
+The Lipschitzian functions are, in general case, multiextremal. If we consider the convolution (\ref{conv}) aggregating multiextremal criteria it will be multiextremal as well.
+
+In the computational scheme proposed in the paper we consider a family $S$ of scalar problems (\ref{conv_problem}) that are solved jointly and take into account the information obtained in the course of already completed optimizations.
+For this purpose in the region of weight coefficients determined by the conditions (\ref{lambda}) a uniform grid is built and the family $S$ consists of the problems (\ref{conv_problem}) with coefficients $\lambda$ corresponding the nodes of the grid.
+
+For solving the problems of the family a global optimization technique based on reducing the multidimensional problem (\ref{conv_problem}) to an equivalent univariate subproblem [Strongin+Sergeyev, Pintér, Zhigljavsky+, Sergeyev+ Strongin, Paulavičius+, Sergeyev+ Kvasov]  in combination with one-dimensional information algorithm with accelerated convergence.
+
+Let us consider this technique in more detail.
+
+To simplify the description of the technique let us present the problems (\ref{conv_problem}) in the following unified form
+\begin{equation}\label{problem_f}
+f(x) \rightarrow \min, \; x \in Q,
+\end{equation}
+where 
+\[
+f(x) = \max_{1 \leq i \leq k}{\lambda_i w_i(x)} + \gamma \sum_{i=1}^k{\lambda_i w_i(x)}
+\]
+and $f(x)$ meets the Lipschitz condition 
+\begin{equation}\label{lip_1}
+\left| f(x) - f(y)\right| \leq L \left\| x-y \right\|, \; x,y, \in Q,
+\end{equation}
+with the constant $L>0$.
+
+It is known as a fundamental fact that $N$-dimensional hyperparallelepiped (\ref{Q}) and the interval $[0,1]$ of the real axis are the equipotent sets and the interval $[0,1]$ can be mapped onto the parallelepiped (\ref{Q}) unambiguously and continuously.
+Such mappings are called \textit{Peano-type curves} or \textit{evolvents}.
+
+Let $x(t), \; t \in [0,1]$  be a Peano-type curve, and the function  $f(x)$ from (\ref{problem_f}) be continuous. As
+\[
+Q =\left\{x(t) : t \in [0,1]\right\},
+\]
+$f(x)$ and $x(t)$ are continuous, then
+\[
+\min_{x \in Q} f(x) = \min_{t \in [0,1]}f(x(t)),
+\]
+i.e., solving the multidimensional problem (\ref{problem_f}) can be reduced to the minimization of the one-dimensional function $f(x(t))$.
+
+However, in the case when the function $f(x)$ is Lipschitzian this property is not valid for the reduced function. The function $f(x(t))$ will satisfy the H{\"o}lder condition
+\begin{equation}\label{holder}
+\left|f(x(t'))-f(x(t''))\right| \leq H \left|t'-t''\right|^{1/N}, \; t',t'' \in [0,1],
+\end{equation}
+with a H{\"o}lder constant $H>0$ \cite{Strongin2000,Sergeyev2013}.
+
+If the function $f(x)$ satisfies the Lipschitz condition (\ref{lip_1}) then $H=2L\sqrt{N+3}$.
+
+As a consequence, minimization of the function $f(x(t))$  requires the use of special algorithms oriented at functions with property (\ref{holder}).
+
+In the proposed scheme for solving reduced univariate minimization problem we apply modified information-statistical algorithm \cite{Strongin2000} with local refinements (ALR) accelerating convergence to global optimum of functions that satisfy the H{\"o}lder condition.
+
+Before description of this algorithm let us follow again the path of simplification and formulate the general one-dimensional optimization problem as
+\begin{equation}\label{problem_fi}
+\varphi(t) \rightarrow \min, \; t \in [\alpha,\beta].
+\end{equation}
+In our consideration $\varphi(t) = f(x(t))$, $[\alpha,\beta] = [0,1]$ and $\varphi(t)$  meets the H{\"o}lder condition with the constant $H$.
+
+The description of ALR can be presented in the following manner.
+
+Let in the optimization problem the term ``trial'' denote the computation of objective function value at a point of the feasible domain.
+
+Solving the first problem from $S$ we reduce it to the univariate problem (\ref{problem_fi}) that is solved by the univariate method ALR.
+This method carries out 2 first iterations at boundary points and the next trials are executed in accordance with the points 1--4 of the algorithm's description up to stopping condition to be met.
+
+In the course of optimization the algorithm builds the trial points $t^1,...,t^n$ which correspond to the points $x^1 = x(t^1),...,x^n=x(t^n)$ in the domain $Q$ .
+At these points the values $v^j = \Phi_\lambda(x^j), \; 1 \leq j \leq n,$ of the function $\Phi_\lambda(x)$ from (\ref{conv_problem}) are computed as values $\varphi(t^j)$ of the function $\varphi_t$  minimized by ALR.
+
+In turn, obtaining the values $\Phi_\lambda(x^j)$  is implemented through the computation of criteria values $a_s^j = w_s(x^j), \; 1 \leq s \leq k, \; 1 \leq j \leq n,$  that can be considered as components of the matrix $A(k \times n)$ containing \textit{a posteriori information} about the problem.
+
+This information can be used for decreasing the number of trials during solving the problem (\ref{conv_problem}) with other weight vector $\lambda$.
+Indeed, if we take the weight vector $\mu = (\mu_1,...,\mu_k) \neq \lambda$, it is easy on the base of matrix $A$ to calculate values $\Phi_\mu(x^j), \; 1 \leq j \leq n,$ at points $x^1 = x(t^1),...,x^n=x(t^n)$, i.e.
+values $\varphi(t^1),...,\varphi(t^n)$ of the function $\varphi(t)$ in the problem (\ref{problem_fi}) being the reduced problem (\ref{conv_problem}) with weight vector $\mu$.
+But in this case in ALR we can take points $t^1,...,t^n$ as points of initial trials in which values $ \varphi(t^j) = \Phi_\mu(x^j), \; 1 \leq j \leq n,$ have been already computed and continue optimization following the rules 1--4 of the algorithm.
+In the course of continuation, ALR will generate new points $t^q$ in which the values $\Phi_\mu(x(t^j))$, and, therefore, values $w_s(x(t^j)), \; 1 \leq s \leq k,$ will be evaluated.
+The latters can be added to the matrix $A$ and be taken into account in the same manner for solving other problems (\ref{conv_problem}).
+
+The proposed scheme of utilizing a posteriori information allows one to decrease significantly the total number of criteria computation when joint optimizing problems of the family $S$.
+On the whole, the given approach is useful  if the time spent to calculate the criteria at one point is significantly higher than the time taken by the algorithm for implementation of the trial at that point, i.e., it is oriented at time consuming MCO problems.
+
+The efficiency of the described general scheme is estimated in the next section in comparison with several evolutionary methods.
 
@@
 \end{document}
++

### 2022-01-06 2022/OM&S/paper.tex
1. section 3

@@
 
+According to ARL applied to solving the problem (\ref{problem_fi}) its first $l \geq 2$  trials are executed at arbitrary points $t^1, \dots, t^l$ of the interval $[\alpha, \beta]$ including points $\alpha$ and $\beta$, and the function values $\nu^1, \dots, \nu^l$ are evaluated at these points, i.e., $\nu^s=\varphi(t^s)$, $1 \leq j \leq l$.
+
+In order to choose a point $t^{n+1}$ of a new $(n+1)$-th trial after completing $n \geq l$ trials it is necessary to implement the following actions.
+
+1. The points $t^1, \dots, t^n$ of completed trials and boundary points of the search interval $[\alpha, \beta]$ are ordered in increasing order and renumbered by subscripts, i.e.,
+
+\begin{equation}\label{eq:15}
+t_1 = \alpha < t_1 < \dots < t_{n-1} < t_n = \beta.
+\end{equation}
+
+The values $\nu_s = \varphi(t_s)$ are juxtaposed to the points  $t_s$, $1 \leq s \leq n$.
+
+2.
+A real values $R(q)$ are assigned to all the intervals $(t_{q-1}, t_q)$, $2 \leq q \leq n$, formed by neighboring points from (\ref{eq:15}).
+$R(q)$ is called \textit{the characteristic} of the interval $(t_{q-1}, t_q)$ and is computed in accordance with the following rule.
+
+If number of completed trials $2 \leq n \leq J$, where $J$ is a predefined number, then 
+\begin{equation}\label{eq:16}
+R(q) = B(q),
+\end{equation}
+otherwise, when $n > J$,
+\begin{equation}\label{eq:17}
+R(q) = B(q)\left( \sqrt{(\nu_q - \varphi^*(n))(\nu_{q-1} - \varphi^*(n))} +M(1.5)^{-\sigma}\right)^{-1}.
+\end{equation}
+
+Here
+\begin{equation}\label{eq:18}
+\begin{matrix}
+B(q) = m (t_s - t_{s-1})^{1/N} + (\nu_q - \nu_{q-1})^2 / (m (t_s - t_{s-1})^{1/N}) -2(\nu_{q-1} + \nu_q), \\
+\varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
+\end{matrix}
+\end{equation}
+is minimal computed value of the objective function $\varphi(t)$ at trial points,
+
+\begin{equation}\label{eq:19}
+M = \max_{2 \leq  s \leq n} { \frac{|\nu_s - \nu_{s-1}|}{(t_s - t_{s-1})^{1/N}} }
+\end{equation}
+
+is underestimation of the H{\"o}lder constant obtained according to the trial results,
+\begin{equation}\label{eq:20}
+m = 
+\begin{cases}
+\begin{matrix}
+rM, & M>0, \\
+1, & M= 0,
+\end{matrix}
+\end{cases}
+\end{equation}
+
+is improved estimation of the H{\"o}lder constant due to method's parameter $r>1$, $\sigma$ is some integer parameter that influences the speed of convergence to the global minimum.
+
+3. Among all intervals, the interval   with the highest characteristic is selected
+\begin{equation}\label{eq:21}
+R(p) = \max_{2 \leq s \leq n} {R(s)}
+\end{equation}
+
+4. The new $(n+1)$-th trial is performed at the point
+\begin{equation}\label{eq:22}
+t^{n+1} = \frac{t_p-t_{p-1}}{2} - sign(\nu_p-\nu_{p-1}) \left( \frac{|\nu_p-\nu_{p-1}|}{M} \right)^N \frac{1}{2r} 
+\end{equation}
+the value $\nu^{n+1} = \varphi(t^{n+1})$ is computed and the iteration number $n$ is increased by 1.
+
+The method ALR described above is a modification of the base information-statistical algorithm [Strongin+Sergeyev, Sergeyev+Strongin] for minimization functions satisfying the H{\"o}lder condition.
+The modification consists in introducing new characteristics (\ref{eq:17}) after initial stage of the search to improve the method's speedup on account of increasing values of characteristics for intervals containing current record (\ref{eq:18}).
+
+Like its prototype, ARL can apply as the stopping condition the inequality
+\begin{equation}\label{eq:23}
+(t_p-t_{p-1})^{1/N} < \varepsilon
+\end{equation}
+i.e., completes the search when the length of the interval with maximal characteristic (\ref{eq:21}) becomes less than the predefined accuracy $\varepsilon > 0$.
+
+The sufficient condition of convergence to global minima in the problem (\ref{problem_fi}) with the objective function under H{\"o}lder property with constant $H$ is
+\begin{equation}\label{eq:24}
+m > 2^{2 - 1 / N}H.
+\end{equation}
+
+If $\varphi(t) = f(x(t)) $ where $x(t)$ is the Peano-type curve and $f(x)$ meets (\ref{lip_1}), then $H=2L\sqrt{N+3}$ and the sufficient condition takes the form 
+\begin{equation}\label{eq:25}
+m > 2^{3 - 1 / N}L\sqrt{N+3}.
+\end{equation}
+
+Finalizing the proposed general scheme  let us describe a procedure  of successive solving the problems of the family $S$ being scalar problems (\ref{conv_problem}) for different weight vectors $\lambda = (\lambda_1, \dots,\lambda_k)$ (\ref{lambda}).
+
 Solving the first problem from $S$ we reduce it to the univariate problem (\ref{problem_fi}) that is solved by the univariate method ALR.

### 2022-01-06 2022/OM&S/paper.tex
Section 3

--- 2022-01-06 2022/OM&S/paper.tex
@@
 
-%\usepackage{color}
+\usepackage{color}
 \usepackage{hyperref}
@@
 
-For solving the problems of the family a global optimization technique based on reducing the multidimensional problem (\ref{conv_problem}) to an equivalent univariate subproblem [Strongin+Sergeyev, Pintér, Zhigljavsky+, Sergeyev+ Strongin, Paulavičius+, Sergeyev+ Kvasov]  in combination with one-dimensional information algorithm with accelerated convergence.
+For solving the problems of the family a global optimization technique based on reducing the multidimensional problem (\ref{conv_problem}) to an equivalent univariate subproblem \cite{Strongin2000,Pinter1996,Zhigljavsky2008,Sergeyev2013,PaulaviciusZilinskas2014,Sergeyev2017} in combination with one-dimensional information algorithm with accelerated convergence.
 
@@
 
+According to ARL applied to solving the problem (\ref{problem_fi}) its first $l \geq 2$  trials are executed at arbitrary points $t^1, \dots, t^l$ of the interval $[\alpha, \beta]$ including points $\alpha$ and $\beta$, and the function values $v^1, \dots, v^l$ are evaluated at these points, i.e., $v^s=\varphi(t^s)$, $1 \leq j \leq l$.
+
+In order to choose a point $t^{n+1}$ of a new $(n+1)$-th trial after completing $n \geq l$ trials it is necessary to implement the following actions.
+
+\begin{enumerate}
+\item 
+The points $t^1, \dots, t^n$ of completed trials and boundary points of the search interval $[\alpha, \beta]$ are ordered in increasing order and renumbered by subscripts, i.e.,
+\begin{equation}\label{eq:15}
+t_1 = \alpha < t_1 < \dots < t_{n-1} < t_n = \beta.
+\end{equation}
+The values $v_s = \varphi(t_s)$ are juxtaposed to the points  $t_s$, $1 \leq s \leq n$.
+\item 
+A real values $R(q)$ are assigned to all the intervals $(t_{q-1}, t_q)$, $2 \leq q \leq n$, formed by neighboring points from (\ref{eq:15}).
+$R(q)$ is called \textit{the characteristic} of the interval $(t_{q-1}, t_q)$ and is computed in accordance with the following rule.
+
+\textcolor[rgb]{1,0,0}{If number of completed trials is not evenly divisible by $J$, where \textit{global-to-local ratio} $J$ is a predefined number, then }
+\begin{equation}\label{eq:16}
+R(q) = B(q),
+\end{equation}
+otherwise
+\begin{equation}\label{eq:17}
+R(q) = B(q)\left( \frac {\sqrt{(v_q - \varphi^*(n))(v_{q-1} - \varphi^*(n))}}{M} +(1.5)^{-\sigma}\right)^{-1}.
+\end{equation}
+Here
+\begin{gather}\label{eq:18}
+B(q) = (t_s - t_{s-1})^{1/N} + \frac{(v_q - v_{q-1})^2} {m^2 (t_s - t_{s-1})^{1/N}} -2\frac{v_{q-1} + v_q}{m},\\
+\varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
+\end{gather}
+is minimal computed value of the objective function $\varphi(t)$ at trial points,
+\begin{equation}\label{eq:19}
+M = \max_{2 \leq  s \leq n} { \frac{|v_s - v_{s-1}|}{(t_s - t_{s-1})^{1/N}} }
+\end{equation}
+is underestimation of the H{\"o}lder constant obtained according to the trial results,
+\begin{equation}\label{eq:20}
+m = 
+\begin{cases}
+\begin{matrix}
+rM, & M>0, \\
+1, & M= 0,
+\end{matrix}
+\end{cases}
+\end{equation}
+is improved estimation of the H{\"o}lder constant due to method's parameter $r>1$, $\sigma$ is some integer parameter that influences the speed of convergence to the global minimum.
+
+\item
+Among all intervals, the interval   with the highest characteristic is selected
+\begin{equation}\label{eq:21}
+R(p) = \max_{2 \leq s \leq n} {R(s)}.
+\end{equation}
+
+\item
+The new $(n+1)$-th trial is performed at the point
+\begin{equation}\label{eq:22}
+t^{n+1} = \frac{t_p-t_{p-1}}{2} - sign(v_p-v_{p-1}) \left( \frac{|v_p-v_{p-1}|}{M} \right)^N \frac{1}{2r} 
+\end{equation}
+the value $v^{n+1} = \varphi(t^{n+1})$ is computed and the iteration number $n$ is increased by 1.
+
+\end{enumerate}
+
+The method ALR described above is a modification of the base information-statistical algorithm \cite{Strongin2000,Sergeyev2013} for minimization functions satisfying the H{\"o}lder condition.
+The modification consists in introducing new characteristics (\ref{eq:17}) after initial stage of the search to improve the method's speedup on account of increasing values of characteristics for intervals containing current record (\ref{eq:18}).
+
+Like its prototype, ARL can apply as the stopping condition the inequality
+\begin{equation}\label{eq:23}
+(t_p-t_{p-1})^{1/N} < \varepsilon,
+\end{equation}
+i.e., completes the search when the length of the interval with maximal characteristic (\ref{eq:21}) becomes less than the predefined accuracy $\varepsilon > 0$.
+
+The sufficient condition of convergence to global minima in the problem (\ref{problem_fi}) with the objective function under H{\"o}lder property with constant $H$ is
+\begin{equation}\label{eq:24}
+m > 2^{2 - 1 / N}H.
+\end{equation}
+
+If $\varphi(t) = f(x(t)) $ where $x(t)$ is the Peano-type curve and $f(x)$ meets (\ref{lip_1}), then $H=2L\sqrt{N+3}$ and the sufficient condition takes the form 
+\begin{equation}\label{eq:25}
+m > 2^{3 - 1 / N}L\sqrt{N+3}.
+\end{equation}
+
+Finalizing the proposed general scheme  let us describe a procedure  of successive solving the problems of the family $S$ being scalar problems (\ref{conv_problem}) for different weight vectors $\lambda = (\lambda_1, \dots,\lambda_k)$ (\ref{lambda}).
+
 Solving the first problem from $S$ we reduce it to the univariate problem (\ref{problem_fi}) that is solved by the univariate method ALR.

### 2022-01-10 2022/OM&S/paper.tex
Corrections in Section 1

@@
 In particular, for the multiextremal case where it is required to find the global optimum the publications \cite{Evtushenko2014,Zilinskas2015,GERGEL2017_1,Gergel2019_2,Barkalov2021} describe applications of the global optimization algorithms to MCO problems in combination with the convolution approach.
+Among these methods of multiextremal optimization the efficient algorithms that utilize  schemes of dimensionality reduction such as Peano mappings \cite{Strongin2000,Sergeyev2013} or nested optimization \cite{Strongin2000,Grishagin2018} are perspective for application in the  frameworks of multiobjective optimization.
 
-Among other methods of solving MCO problems it is necessary to pay attention to the important class of metaheuristic nature-inspired algorithms.
+As for other methods of solving MCO problems it is necessary to pay attention to the important class of metaheuristic nature-inspired algorithms.
 Many of them are nature-inspired, i.e.,  are  based on modeling physical processes (for example, simulated annealing \cite{Locatelli2002,Aarts2014}) or mainly behavior of biological agents including evolutionary \cite{Price2005,Coello2007} and as their part genetic algorithms \cite{Ruiz2015,Deb2002,Zitzler2001} and particle swarm methods modeling the swarming intelligence of bees, fireflies, birds, ants, etc.
@@
 
-\textcolor[rgb]{1,0,0}{If number of completed trials is not evenly divisible by $J$, where \textit{global-to-local ratio} $J$ is a predefined number, then }
+If number of completed trials is not exactly divisible by $J$, where \textit{global-to-local ratio} $J$ is a predefined number, then
 \begin{equation}\label{eq:16}
@@
 
-Finalizing the proposed general scheme  let us describe a procedure  of successive solving the problems of the family $S$ being scalar problems (\ref{conv_problem}) for different weight vectors $\lambda = (\lambda_1, \dots,\lambda_k)$ (\ref{lambda}).
+Finalizing the proposed general scheme  let us describe a procedure  of successive solving the problems of the family $S$ which are scalar problems (\ref{conv_problem}) for different weight vectors $\lambda = (\lambda_1, \dots,\lambda_k)$ (\ref{lambda}).
+This procedure will be called Repeated Algorithm with Local Refinements and Data Accumulation (RALR-DA).
 

### 2022-01-10 2022/OM&S/paper.tex
Begining of Section 4

@@
 
-\section{Numerical experiments}
+\section{Computational experiments}
+
+For efficiency assessment of the computational scheme described in the previous section the experiment with multiextremal multidimensional MCO benchmarks was conducted.
+These benchmarks were constructed on the base of multiextremal functions belonging to the class GKLS \cite{Gaviano2003} widely used in testing global optimization methods, i.e., the partial criteria $w_i(x), \; 1 \leq i \leq k,$ were generated as GKLS functions.
+
+Along with the proposed technique these MCO benchmarks were solved by 5 metaheuristic methods.
+Two algorithms -- SMPSO \cite{Nebro2009} and OMOPSO \cite{Durillo2010} -- belong to the class of particle swarm methods and 3 others are genetic ones (NSGA-II \cite{Deb2002}, SPEA2 \cite{Zitzler2001}, IBEA \cite{Zitzler2004}).
+Implementation of these methods was taken from the library jMetalPy \cite{Benitez2019}.
+It should be noted that metaheuristic algorithms do not use convolutions to assess the Pareto set.
+
+For estimating the effectiveness of solving the test MCO problem the hypervolume (HV) index  \cite{} was used as the quality criterion of the Pareto set approximations built by the methods.
+Moreover, besides comparing the efficiency of tested algorithms via their HVs, an ``ideal'' value of HV was computed by means of selection of non-dominated nodes of a dense uniform grid placed in the feasible domain.
+
+In total, three groups of experiments have been performed.
+Each group contained a series of MCO problems of the same dimension and number of criteria.
+Since all GKLS functions have a global minimum value equal to $-1$, the partial  criteria $w_i(x), \; 1 \leq i \leq k,$  increased by a positive constant to ensure the requirements (\ref{positive}).
+
+In all experiments the domain
+\[
+Q = \left\{x \in R^N : -1 \leq x_j \leq -1, 1 \leq i \leq N \right\}
+\]
+was considered as the feasible one.
+
+Each problem was solved by all the methods including the grid technique and corresponding HVs were determined.
+Metaheuristic methods because of their stochastic nature sometimes realize unsuccessful runs.
+In order to estimate better their efficiency each of these methods carried out 10 launches when solving a separate MCO problem and the average result was taken into account.
+
+During all experiments the method ALR was used with parameters $r=5.6$, $\sigma = 15$, $T=3$.
+Accuracy $\varepsilon$ from the stopping condition (\ref{eq:23}) was $0.05$ in bi-criteria problems and $0.01$ in 4-criteria ones.
+
+For correct comparison of the algorithms (except grid method, of course) their parameters influencing the stopping condition were chosen to provide approximately the same number of trials.
+After the completion of solving all the group problems, HVs of each method were averaged.
+
+In the first experiment 100 two-dimensional bi-criteria problems have been taken.
+The contour plots of a convolution $\Phi_\lambda(x)$ from (\ref{conv_problem})  demonstrating its multiextremal and non-smooth relief are presented in Fig.~\ref{fig1}.
+For estimating the Pareto set by RALR-DA, in each MCO problem 50 convolutions (\ref{conv_problem}) corresponding to 50 values $\lambda$ uniformly distributed over interval $[0,1]$ have been minimized.
+
+The Pareto front in one of test problems is shown in Fig.~\ref{fig2}.
+
+The results of the experiment are presented in Tab.~\ref{tab1} and Fig.~\ref{fig3} where Pareto corresponds to the estimation obtained by the grid method.
 

### 2022-01-10 2022/OM&S/paper.tex
Section 4

@@
 In particular, for the multiextremal case where it is required to find the global optimum the publications \cite{Evtushenko2014,Zilinskas2015,GERGEL2017_1,Gergel2019_2,Barkalov2021} describe applications of the global optimization algorithms to MCO problems in combination with the convolution approach.
-Among these methods of multiextremal optimization the efficient algorithms that utilize  schemes of dimensionality reduction such as Peano mappings \cite{Strongin2000,Sergeyev2013} or nested optimization \cite{Strongin2000,Grishagin2018} are perspective for application in the  frameworks of multiobjective optimization.
+Among these methods of multiextremal optimization the efficient algorithms that utilize  schemes of dimensionality reduction such as Peano mappings \cite{Strongin2000,Sergeyev2013} or nested optimization \cite{Strongin2000,Grishagin2015_2,Grishagin2018} are perspective for application in the  frameworks of multiobjective optimization.
 
@@
 
-If number of completed trials is not exactly divisible by $J$, where \textit{global-to-local ratio} $J$ is a predefined number, then
+If number of completed trials is not exactly divisible by $T$, where \textit{global-to-local ratio} $T$ is a predefined number, then
 \begin{equation}\label{eq:16}
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{fig1.png} 
+\caption{Contour plots of 2-dimensional convolution.}\label{fig1} 
+\end{figure}
+
 The Pareto front in one of test problems is shown in Fig.~\ref{fig2}.
 
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{fig2.png} 
+\caption{Contour plots of 2-dimensional convolution.}\label{fig2} 
+\end{figure}
+
 The results of the experiment are presented in Tab.~\ref{tab1} and Fig.~\ref{fig3} where Pareto corresponds to the estimation obtained by the grid method.
+
+\begin{table}
+\tbl{Results of the experiment with bi-criteria MCO problems of dimension 2.}
+{\begin{tabular}{ccccccc} \toprule
+ Pareto & RALR-DA & OMOPSO & SMPSO & NSGAII & IBEA & SPEA2 \\ \midrule
+  6.62 & 6.44 & 5.93 & 5.88 & 5.71 & 5.65 & 5.61 \\ \bottomrule
+\end{tabular}}
+\label{tab1}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{fig3.pdf} 
+\caption{HV indices in bi-criteria 2-dimensional case.}\label{fig3} 
+\end{figure}
+
+The second experiment was aimed at solving bi-criteria 4-dimensional MCO problems.
+10 test problems have been investigated and for each of them, 100 convolutions (\ref{conv_problem}) with uniformly distributed weight coefficients $\lambda$ were minimized by RALR-DA.
+Tab.~\ref{tab2} contains the results achieved by the methods participated in the experiment and Fig.~\ref{fig4} presents them in graphical form.
+
+\begin{table}
+\tbl{Results of testing bi-criteria 4-dimensional MCO problems.}
+{\begin{tabular}{ccccccc} \toprule
+ Pareto & RALR-DA & SPEA2 & OMOPSO & IBEA & NSGAII & SMPSO   \\ \midrule
+  5.04 & 4.71 & 4.53 & 4.45 & 4.39 & 4.39 & 4.37 \\ \bottomrule
+\end{tabular}}
+\label{tab2}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{fig4.pdf} 
+\caption{HV indices in bi-criteria 4-dimensional case.}\label{fig4} 
+\end{figure}
+
+The third collection of test problems consisted of 10 two-dimensional MCO benchmarks with 4 criteria.
+In this experiment RALR-DA in each MCO problem minimized convolutions (\ref{conv_problem}) corresponding to 100 different values of weight coefficients $\lambda$ uniformly distributed in the domain (\ref{lambda}).
+Like  the  previous experiments, the results are presented in Tab.~\ref{tab3} and Fig.~\ref{fig5}.
+
+\begin{table}
+\tbl{Results of testing bi-criteria 4-dimensional MCO problems.}
+{\begin{tabular}{ccccccc} \toprule
+ Pareto & RALR-DA & OMOPSO & SPEA2 & SMPSO & IBEA & NSGAII   \\ \midrule
+  23.89 & 23.52 & 22.87 & 22.80 & 22.53 & 20.68 & 19.82 \\ \bottomrule
+\end{tabular}}
+\label{tab3}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{fig5.pdf} 
+\caption{HV indices in 4-criteria 2-dimensional case.}\label{fig5} 
+\end{figure}
+
+As it can be seen from the presented experiments, the proposed method demonstrates the best results among all methods involved in evaluation of their effectiveness based on HV index.
+Moreover, the efficiency of RALR-DA is close to the best one estimated on a dense uniform grid.
+
+As for metaheuristic algorithms, their results differ in different experiments.
+For example, in 4-dimensinal bi-criteria case the method SPEA is the best among nature-inspired algorithms, but it is the worst in 2-dimensional problems with 2 criteria.
+On the whole, as the most stable metaheuristic method, the algorithm OMOPSO can be recognized.
 

### 2022-01-11 2022/OM&S/paper.tex
add one reference

@@
 
-For estimating the effectiveness of solving the test MCO problem the hypervolume (HV) index  \cite{} was used as the quality criterion of the Pareto set approximations built by the methods.
+For estimating the effectiveness of solving the test MCO problem the hypervolume (HV) index  \cite{Evtushenko2014} was used as the quality criterion of the Pareto set approximations built by the methods.
 Moreover, besides comparing the efficiency of tested algorithms via their HVs, an ``ideal'' value of HV was computed by means of selection of non-dominated nodes of a dense uniform grid placed in the feasible domain.
@@
 
-During all experiments the method ALR was used with parameters $r=5.6$, $\sigma = 15$, $T=3$.
+During all experiments the method ALR was used with parameters $r=5.6$, $\sigma = 18$, $T=3$. 
+\textcolor[rgb]{1,0.41,0.13}{The use of $T=3$ corresponds to alternation of one ``local'' iteration and two ``global'' iterations.}
 Accuracy $\varepsilon$ from the stopping condition (\ref{eq:23}) was $0.05$ in bi-criteria problems and $0.01$ in 4-criteria ones.

### 2022-01-13 2022/OM&S/paper.tex
abstract

@@
 \begin{abstract}
-This template is for authors who are preparing a manuscript for a Taylor \& Francis journal using the \LaTeX\ document preparation system and the \texttt{interact} class file, which is available via selected journals' home pages on the Taylor \& Francis website.
+The paper is devoted to consideration of multiextremal black-box MCO problems with time-consuming criteria and proposes a new computational algorithm for solving such problems.
+This algorithm is based on combination of convolution approach, dimensionality reduction, and information-statistical global optimization algorithm.
+For approximating the Pareto set a family of maximum convolutions is considered generating problems of scalar multiextremal optimization.
+The latters are reduced by means of Peano mappings to the equivalent problems of univariate optimization that are solved by a modification of the information-statistical global optimization method.
+This method applies the technique of local refinements and uses in the course of current univariate optimization the information about criteria that was obtained when solving completed optimization problems.
+These features allow one to accelerate the search of global solution.
+The efficiency of the proposed MCO algorithm is experimentally estimated on representative classes of multidimensional multiextremal test problems.
+Comparison with 5 nature-inspired particle swarm and genetic algorithms on the base of the hypervolume index reflecting quality of Pareto set evaluation demonstrates qualitative results of the proposed method.
 \end{abstract}
@@
 
-In general case, there exist three key factors determining the complexity of analyzing these models.
+In general case, there exist three key factors that determine the complexity of analyzing these models.
 The crucial feature corresponds to the typical case where criteria are contradictory: increasing one criterion leads to decreasing the other.
@@
 
-In the next section we describe an information-statistical algorithm with accelerated convergence and its application to searching for Pareto optimal solutions in multiextremal MCO problems.
+In the next section propose a modified information-statistical algorithm with accelerated convergence and its application to searching for Pareto optimal solutions in multiextremal MCO problems.
 
@@
 
-The Lipschitzian functions are, in general case, multiextremal. If we consider the convolution (\ref{conv}) aggregating multiextremal criteria it will be multiextremal as well.
+The Lipschitzian functions are, in general case, multiextremal. If we consider the convolution (\ref{conv}) aggregating multiextremal criteria it is multiextremal as well.
 
 In the computational scheme proposed in the paper we consider a family $S$ of scalar problems (\ref{conv_problem}) that are solved jointly and take into account the information obtained in the course of already completed optimizations.
-For this purpose in the region of weight coefficients determined by the conditions (\ref{lambda}) a uniform grid is built and the family $S$ consists of the problems (\ref{conv_problem}) with coefficients $\lambda$ corresponding the nodes of the grid.
+For this purpose in the region of weight coefficients determined by the conditions (\ref{lambda}) a uniform grid is built and the family $S$ consists of the problems (\ref{conv_problem}) with coefficients $\lambda$ corresponding to the grid nodes.
 
@@
 Finalizing the proposed general scheme  let us describe a procedure  of successive solving the problems of the family $S$ which are scalar problems (\ref{conv_problem}) for different weight vectors $\lambda = (\lambda_1, \dots,\lambda_k)$ (\ref{lambda}).
-This procedure will be called Repeated Algorithm with Local Refinements and Data Accumulation (RALR-DA).
+This procedure will be called Reduced Algorithm with Local Refinements and Data Accumulation (RALR-DA).
 
@@
 During all experiments the method ALR was used with parameters $r=5.6$, $\sigma = 18$, $T=3$. 
-\textcolor[rgb]{1,0.41,0.13}{The use of $T=3$ corresponds to alternation of one ``local'' iteration and two ``global'' iterations.}
+\textcolor[rgb]{1,0,0}{The use of $T=3$ corresponds to alternation of one ``local'' iteration and two ``global'' iterations.}
 Accuracy $\varepsilon$ from the stopping condition (\ref{eq:23}) was $0.05$ in bi-criteria problems and $0.01$ in 4-criteria ones.
@@
 \centering
-\includegraphics[width=0.6\textwidth]{fig1.png} 
+\includegraphics[width=0.7\textwidth]{fig1.png} 
 \caption{Contour plots of 2-dimensional convolution.}\label{fig1} 
@@
 
-The Pareto front in one of test problems is shown in Fig.~\ref{fig2}.
+\textcolor[rgb]{1,0,0}{The Pareto front in one of test problems is shown in Fig.~\ref{fig2a} and Fig.~\ref{fig2b}.
+Here Fig.~\ref{fig2a} corresponds the Pareto front computed on the uniform grid with the step 0.05 in the search domain whereas Fig.~\ref{fig2b} shows the Pareto front computed by RALR-DA.
+In the last case it is required only 350 trials to obtain the solution with the same accuracy.}
 
@@
 \centering
-\includegraphics[width=0.7\textwidth]{fig2.png} 
-\caption{Contour plots of 2-dimensional convolution.}\label{fig2} 
+\includegraphics[width=0.7\textwidth]{fig2a.png} 
+\caption{Pareto front computed by grid technique.}\label{fig2a} 
 \end{figure}
 
-The results of the experiment are presented in Tab.~\ref{tab1} and Fig.~\ref{fig3} where Pareto corresponds to the estimation obtained by the grid method.
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{fig2b.png} 
+\caption{Pareto front computed by RALR-DA.}\label{fig2b} 
+\end{figure}
+
+\textcolor[rgb]{1,0,0}{The averaged HV indices obtained while solving the two-dimensional bi-criteria problems are presented} in Tab.~\ref{tab1} and Fig.~\ref{fig3} where ``Pareto'' corresponds to the estimation obtained by the grid method.
 
@@
 As for metaheuristic algorithms, their results differ in different experiments.
-For example, in 4-dimensinal bi-criteria case the method SPEA is the best among nature-inspired algorithms, but it is the worst in 2-dimensional problems with 2 criteria.
+For example, in 4-dimensional bi-criteria case the method SPEA is the best among nature-inspired algorithms, but it is the worst in 2-dimensional problems with 2 criteria.
 On the whole, as the most stable metaheuristic method, the algorithm OMOPSO can be recognized.

### 2022-01-13 2022/OM&S/paper.tex
almost done

@@
 
-\title{A fast algorithm for finding efficient decisions in multi-objective problems with black-box multiextremal functions}
+\title{A fast algorithm for finding efficient solutions in multi-objective problems with black-box multiextremal criteria}
 
 \author{
-\name{Konstantin Barkalov, Vladimir Grishagin and Evgeny Kozinov\thanks{CONTACT Evgeny Kozinov. Email: evgeny.kozinov@itmm.unn.ru}}
+\name{Konstantin Barkalov, Vladimir Grishagin and Evgeny Kozinov\thanks{CONTACT Konstantin Barkalov. Email: konstantin.barkalov@itmm.unn.ru}}
 \affil{Software Department, Lobachevsky State University, Nizhni Novgorod, Russia}
@@
 
-The method ALR described above is a modification of the base information-statistical algorithm \cite{Strongin2000,Sergeyev2013} for minimization functions satisfying the H{\"o}lder condition.
+The method ALR described above is a modification of the basic information-statistical algorithm \cite{Strongin2000,Sergeyev2013} for minimization functions satisfying the H{\"o}lder condition.
 The modification consists in introducing new characteristics (\ref{eq:17}) after initial stage of the search to improve the method's speedup on account of increasing values of characteristics for intervals containing current record (\ref{eq:18}).
+\textcolor[rgb]{1,0,0}{Actually, the second term in (\ref{eq:17}) will have the largest value equal to $(1.5)^{\sigma}$ only for those intervals for which the equality $v_q = \varphi^*(n)$ or $v_{q-1} = \varphi^*(n)$ holds, i.e.
+for the intervals with the boundary points correspond to the current record $\varphi^*(n)$.}
 
@@
 \end{equation}
+
+\textcolor[rgb]{1,0,0}{The complete convergence theory of the information-statistical algorithm is given in \cite{Strongin2000,Sergeyev2013}.}
 
@@
 During all experiments the method ALR was used with parameters $r=5.6$, $\sigma = 18$, $T=3$. 
-\textcolor[rgb]{1,0,0}{The use of $T=3$ corresponds to alternation of one ``local'' iteration and two ``global'' iterations.}
+\textcolor[rgb]{1,0,0}{The use of $T=3$ corresponds to alternation of one ``local'' iteration and two ``global'' iterations, that can be interpreted as a combination of local refinement and global updating of the current optimizer.}
 Accuracy $\varepsilon$ from the stopping condition (\ref{eq:23}) was $0.05$ in bi-criteria problems and $0.01$ in 4-criteria ones.
@@
 \textcolor[rgb]{1,0,0}{The Pareto front in one of test problems is shown in Fig.~\ref{fig2a} and Fig.~\ref{fig2b}.
-Here Fig.~\ref{fig2a} corresponds the Pareto front computed on the uniform grid with the step 0.05 in the search domain whereas Fig.~\ref{fig2b} shows the Pareto front computed by RALR-DA.
-In the last case it is required only 350 trials to obtain the solution with the same accuracy.}
+Here Fig.~\ref{fig2a} corresponds to the Pareto front computed on the uniform grid with the step 0.05 in the search domain whereas Fig.~\ref{fig2b} shows the Pareto front computed by RALR-DA.
+In the last case it is required only 608 trials to obtain the solution with the same accuracy.}
 
@@
 
+\section{Conclusion}
+
+In the paper the multidimensional problems of multicriteria (multiobjective) optimization with black-box time-consuming criteria satisfying the Lipschitz condition are considered.
+These assumptions lead to significant complexity of the problems under investigation and require the development of special efficient numerical methods for their solving.
+For this goal in the paper a new algorithm is proposed and its computational scheme is described.
+The algorithm is based on reducing the MCO problem to a family of single-criterion (scalar) optimization problems generating Pareto-optimal solutions via building parameterized convolutions.
+In turn, each multidimensional scalar problem is reduced to a univariate one by means of Peano curves that map the multidimensional feasible domain into the real axis.
+Under Lipschitz condition for criteria the objective functions of the reduced one-dimensional problems satisfy the H{\"o}lder condition and for solving these problems it is necessary to apply specialized global optimization methods.
+As such method we propose a new algorithm that combines in its decision rule iteration with ``local'' and ``global'' goals.
+Global iterations provide the convergence to the global optimum and the local ones accelerate the search.
+Moreover, while solving the univariate problems generated by the family of convolutions the proposed algorithm is able to take into account the criteria values computed already in the course of solving the univariate problems completed earlier.
+This feature reduces significantly number of trials (calculations of criteria values): the later a problem is solved, the less new criteria computations are produced.
+
+The quality of the proposed algorithm is studied in experiment on 3 representative sets of multiextremal MCO problems of different dimensions and criteria number in comparison with 5 particle swarm and genetic methods.
+As a criterion of methods' efficiency the hypervolume index estimating the quality of Pareto set approximation is taken.
+Among all competitors the proposed algorithm demonstrates the best results.
+
+As the directions of further development of the current research it would be interesting to investigate the possibilities of applying other schemes of dimensionality reduction, for example,
+\textcolor[rgb]{1,0,0}{the nested optimization scheme \cite{Grishagin2016,Grishagin2018}, the diagonal \cite{Sergeyev2015,Sergeyev2017} or the simplicial \cite{Zilinskas2011,PaulaviciusZilinskas2014} partitions methods}
+in the framework of the general approach used in the paper. Moreover, the prospective way is to develop parallel versions of the proposed general scheme oriented at powerful supercomputer systems.
+
 \section*{Disclosure statement}

### 2022-01-14 2022/OM&S/paper.tex
almost done

@@
 
-\title{A fast algorithm for finding efficient solutions in multi-objective problems with black-box multiextremal criteria}
+\title{An accelerated algorithm for finding efficient solutions in multiobjective problems with black-box multiextremal criteria}
 
@@
 \name{Konstantin Barkalov, Vladimir Grishagin and Evgeny Kozinov\thanks{CONTACT Konstantin Barkalov. Email: konstantin.barkalov@itmm.unn.ru}}
-\affil{Software Department, Lobachevsky State University, Nizhni Novgorod, Russia}
+\affil{Institute of Information Technology, Mathematics and Mechanics, Lobachevsky State University, 23 Gagarin ave., 603022 Nizhny Novgorod, Russia}
 }
@@
 \begin{abstract}
-The paper is devoted to consideration of multiextremal black-box MCO problems with time-consuming criteria and proposes a new computational algorithm for solving such problems.
+The paper is devoted to consideration of multiextremal multicriteria black-box problems with time-consuming criteria and proposes a new computational algorithm for solving such problems.
 This algorithm is based on combination of convolution approach, dimensionality reduction, and information-statistical global optimization algorithm.
 For approximating the Pareto set a family of maximum convolutions is considered generating problems of scalar multiextremal optimization.
-The latters are reduced by means of Peano mappings to the equivalent problems of univariate optimization that are solved by a modification of the information-statistical global optimization method.
+The latters are reduced by means of Peano mappings to the equivalent problems of univariate optimization that are solved by an accelerated information-statistical global optimization method.
 This method applies the technique of local refinements and uses in the course of current univariate optimization the information about criteria that was obtained when solving completed optimization problems.
 These features allow one to accelerate the search of global solution.
-The efficiency of the proposed MCO algorithm is experimentally estimated on representative classes of multidimensional multiextremal test problems.
+The efficiency of the proposed algorithm is experimentally estimated on representative classes of multidimensional multiextremal test problems.
 Comparison with 5 nature-inspired particle swarm and genetic algorithms on the base of the hypervolume index reflecting quality of Pareto set evaluation demonstrates qualitative results of the proposed method.
@@
 
-In the field of mathematical modeling the intelligent decision-making processes, the problems of multicriterial optimization (MCO) present a wide class of models that are rich and interesting from a theoretical point of view and very important in real applications \cite{Marler2009,Hillermeier2005}.
-Multicriterial problems describe decision-making situations with multiple goals which are considered as functional criteria to be optimized.
+In the field of mathematical modeling the intelligent decision-making processes, the problems of multicriteria optimization (MCO) present a wide class of models that are rich and interesting from a theoretical point of view and very important in real applications \cite{Marler2009,Hillermeier2005}.
+Multicriteria problems describe decision-making situations with multiple goals which are considered as functional criteria to be optimized.
 
@@
 Moreover, the combination of such factors as multiextremality and greater dimension leads to significant computational complexity of the problem because the computational cost increases exponentially when dimension grows.
-From this point of view the multiextremal problems are the most complicated ones in multicriterial optimization and this MCO class is the subject of interest in the given paper.
+From this point of view the multiextremal problems are the most complicated ones in multicriteria optimization and this MCO class is the subject of interest in the given paper.
 
@@
 In this scheme a non-negative coefficient (weight) is assigned to each criterion which reflects numerically an ``importance'' of the criterion and on the base of such weighted criteria a scalar function (convolution) is built and minimized.
-Under appropriate choice of convolution the solution of the built scalar optimization problem is a partial solution of the initial multicriterial problem.
+Under appropriate choice of convolution the solution of the built scalar optimization problem is a partial solution of the initial multicriteria problem.
 Changing weights we can obtain different Pareto points.
@@
 
-The problem of multicriterial optimization is formulated as minimizing over the domain (\ref{Q}) the vector function
+The problem of multicriteria optimization is formulated as minimizing over the domain (\ref{Q}) the vector function
 \begin{equation}\label{W}
@@
 \end{equation}
-
 If
@@
 \end{equation}
-with a sufficiently small parameter $\gamma > 0$ is a Pareto optimal point of the initial MCO problem (\ref{problem}) \cite{Wierzbicki} [Wierzbicki, Krasnoshekov].
+with a sufficiently small parameter $\gamma > 0$ is a Pareto optimal point of the initial MCO problem (\ref{problem}) \cite{Wierzbicki,Marler2004}.
 The positiveness of criteria is not restrictive requirement because it is possible to transform easily the MCO problem with non-positive criteria to the form (\ref{positive}) without loss of Pareto solution.
@@
 If the function $f(x)$ satisfies the Lipschitz condition (\ref{lip_1}) then $H=2L\sqrt{N+3}$.
-
 As a consequence, minimization of the function $f(x(t))$  requires the use of special algorithms oriented at functions with property (\ref{holder}).
 
-In the proposed scheme for solving reduced univariate minimization problem we apply modified information-statistical algorithm \cite{Strongin2000} with local refinements (ALR) accelerating convergence to global optimum of functions that satisfy the H{\"o}lder condition.
+In the proposed scheme for solving reduced univariate minimization problem we apply modified information-statistical Algorithm \cite{Strongin2000} with Local Refinements (ALR) accelerating convergence to global optimum of functions that satisfy the H{\"o}lder condition.
 
@@
 \begin{equation}\label{eq:22}
-t^{n+1} = \frac{t_p-t_{p-1}}{2} - sign(v_p-v_{p-1}) \left( \frac{|v_p-v_{p-1}|}{M} \right)^N \frac{1}{2r} 
+t^{n+1} = \frac{t_p-t_{p-1}}{2} - sign(v_p-v_{p-1}) \left( \frac{|v_p-v_{p-1}|}{M} \right)^N \frac{1}{2r},
 \end{equation}
@@
 The modification consists in introducing new characteristics (\ref{eq:17}) after initial stage of the search to improve the method's speedup on account of increasing values of characteristics for intervals containing current record (\ref{eq:18}).
-\textcolor[rgb]{1,0,0}{Actually, the second term in (\ref{eq:17}) will have the largest value equal to $(1.5)^{\sigma}$ only for those intervals for which the equality $v_q = \varphi^*(n)$ or $v_{q-1} = \varphi^*(n)$ holds, i.e.
-for the intervals with the boundary points correspond to the current record $\varphi^*(n)$.}
+Actually, the second term in (\ref{eq:17}) will have the largest value equal to $(1.5)^{\sigma}$ only for those intervals for which the equality $v_q = \varphi^*(n)$ or $v_{q-1} = \varphi^*(n)$ holds, i.e., for the intervals with the boundary points one of which coincides with the current record $\varphi^*(n)$.
 
@@
 
-\textcolor[rgb]{1,0,0}{The complete convergence theory of the information-statistical algorithm is given in \cite{Strongin2000,Sergeyev2013}.}
+The complete convergence theory of the information-statistical algorithms is given in \cite{Strongin2000,Sergeyev2013}.
 
@@
 During all experiments the method ALR was used with parameters $r=5.6$, $\sigma = 18$, $T=3$. 
-\textcolor[rgb]{1,0,0}{The use of $T=3$ corresponds to alternation of one ``local'' iteration and two ``global'' iterations, that can be interpreted as a combination of local refinement and global updating of the current optimizer.}
+The use of $T=3$ corresponds to alternation of one ``local'' iteration and two ``global'' iterations, that can be interpreted as a combination of local refinement and global updating of the current optimizer.
 Accuracy $\varepsilon$ from the stopping condition (\ref{eq:23}) was $0.05$ in bi-criteria problems and $0.01$ in 4-criteria ones.
@@
 
-\textcolor[rgb]{1,0,0}{The Pareto front in one of test problems is shown in Fig.~\ref{fig2a} and Fig.~\ref{fig2b}.
+The Pareto front in one of test problems is shown in Fig.~\ref{fig2a} and Fig.~\ref{fig2b}.
 Here Fig.~\ref{fig2a} corresponds to the Pareto front computed on the uniform grid with the step 0.05 in the search domain whereas Fig.~\ref{fig2b} shows the Pareto front computed by RALR-DA.
-In the last case it is required only 608 trials to obtain the solution with the same accuracy.}
 
@@
 
-\textcolor[rgb]{1,0,0}{The averaged HV indices obtained while solving the two-dimensional bi-criteria problems are presented} in Tab.~\ref{tab1} and Fig.~\ref{fig3} where ``Pareto'' corresponds to the estimation obtained by the grid method.
+The averaged HV indices obtained while solving the two-dimensional bi-criteria problems are presented in Tab.~\ref{tab1} and Fig.~\ref{fig3} where ``Pareto'' corresponds to the estimation obtained by the grid method.
 
@@
 \begin{table}
-\tbl{Results of testing bi-criteria 4-dimensional MCO problems.}
+\tbl{Results of testing 2-dimensional MCO problems with 4 criteria.}
 {\begin{tabular}{ccccccc} \toprule
@@
 As the directions of further development of the current research it would be interesting to investigate the possibilities of applying other schemes of dimensionality reduction, for example,
-\textcolor[rgb]{1,0,0}{the nested optimization scheme \cite{Grishagin2016,Grishagin2018}, the diagonal \cite{Sergeyev2015,Sergeyev2017} or the simplicial \cite{Zilinskas2011,PaulaviciusZilinskas2014} partitions methods}
+the nested optimization scheme \cite{Grishagin2016,Grishagin2018}, the diagonal \cite{Sergeyev2015,Sergeyev2017} or simplicial \cite{Zilinskas2011,PaulaviciusZilinskas2014} partition methods
 in the framework of the general approach used in the paper. Moreover, the prospective way is to develop parallel versions of the proposed general scheme oriented at powerful supercomputer systems.

### 2022-01-14 2022/OM&S/paper.tex
before submission

@@
 \begin{keywords}
-Multi-objective optimization; global optimization; dimensionality reduction; search information; nature-inspired metaheuristics; numerical comparison
+Multiobjective optimization; global optimization; dimensionality reduction; search information; nature-inspired metaheuristics; numerical comparison
 \end{keywords}
@@
 
-The simplest way to build an approximation of the Pareto set consist in immersion in the feasible domain of problem's parameters a uniform grid (deterministic or random) and selection of all non-dominated nodes of the grid.
+The simplest way to build an approximation of the Pareto set consists in immersion in the feasible domain of problem's parameters a uniform grid (deterministic or random) and in selection of all non-dominated nodes of the grid.
 Unfortunately, this universal approach is too costly because it requires in the worst case pairwise comparison of all the grid nodes and does not take into account the properties of the problem for reducing the computational costs.
@@
 
-As for other methods of solving MCO problems it is necessary to pay attention to the important class of metaheuristic nature-inspired algorithms.
+Note that special classes of multiextremal optimization problems arise when the criteria can be described analytically.
+In this case one can apply efficient methods based on d.c.
+optimization approach \cite{Strekalovsky2001,Strekalovsky2003,Strekalovsky2020}.
+However, this paper considers a completely different case, when the analytical expressions for the functions are not known.
+This situation is typical for finding the best decisions in applied problems in which the evaluation of criteria requires numerical simulation.
+
+As for other methods of solving MCO problems it is necessary to pay attention to the important class of metaheuristic algorithms.
 Many of them are nature-inspired, i.e.,  are  based on modeling physical processes (for example, simulated annealing \cite{Locatelli2002,Aarts2014}) or mainly behavior of biological agents including evolutionary \cite{Price2005,Coello2007} and as their part genetic algorithms \cite{Ruiz2015,Deb2002,Zitzler2001} and particle swarm methods modeling the swarming intelligence of bees, fireflies, birds, ants, etc.
@@
 In this paper we consider multiextremal black-box MCO problems and propose a new scheme of solving these problems.
-This scheme is based on maximum convolution, dimensionality reduction by means of Peano curves, information-statistical global optimization algorithm used for the first time in MCO problems.
+This scheme is based on maximum convolution, dimensionality reduction by means of Peano curves, modified information-statistical global optimization algorithm used for the first time in MCO problems.
 Moreover, in the framework of the scheme the information already obtained in the course of solving scalar subproblems of convolution optimization can be utilized for solving a new convolution subproblem.
 Briefly, the way is as follows.
-Before beginning the optimization of a new convolution we have at our disposal values of criteria computed already and now can recalculate them to values of the current convolution taking into account these values as initial information for optimization algorithm.
+Before beginning the optimization of a new convolution we have at our disposal values of criteria computed earlier and now can recalculate them to values of the current convolution taking into account these values as initial information for optimization algorithm.
 This procedure accelerates significantly the optimization process.
@@
 
-In the next section propose a modified information-statistical algorithm with accelerated convergence and its application to searching for Pareto optimal solutions in multiextremal MCO problems.
+In the next section we propose a modified information-statistical algorithm with accelerated convergence and its application to searching for Pareto optimal solutions in multiextremal MCO problems.
 
@@
 
-In this section we consider the MCO problem under additional assumption according to which the criteria $w_i(x), \; 1 \leq i \leq k,$ are black-box functions and satisfy in the domain $Q$ the Lipschitz condition
+In this section we consider the MCO problem under additional assumptions according to which the criteria $w_i(x), \; 1 \leq i \leq k,$ are black-box functions and satisfy in the domain $Q$ the Lipschitz condition
 \begin{equation}\label{lip}
@@
 
-For solving the problems of the family a global optimization technique based on reducing the multidimensional problem (\ref{conv_problem}) to an equivalent univariate subproblem \cite{Strongin2000,Pinter1996,Zhigljavsky2008,Sergeyev2013,PaulaviciusZilinskas2014,Sergeyev2017} in combination with one-dimensional information algorithm with accelerated convergence.
+For solving the problems of the family, a global optimization technique based on reducing the multidimensional problem (\ref{conv_problem}) to an equivalent univariate subproblem \cite{Strongin2000,Pinter1996,Zhigljavsky2008,Sergeyev2013,PaulaviciusZilinskas2014,Sergeyev2017} in combination with one-dimensional information algorithm with accelerated convergence is proposed.
 
@@
 
-To simplify the description of the technique let us present the problems (\ref{conv_problem}) in the following unified form
+To simplify the description let us present the problems (\ref{conv_problem}) in the following unified form
 \begin{equation}\label{problem_f}
@@
 \item 
-The points $t^1, \dots, t^n$ of completed trials and boundary points of the search interval $[\alpha, \beta]$ are ordered in increasing order and renumbered by subscripts, i.e.,
+The points $t^1, \dots, t^n$ of completed trials are ordered in increasing order and renumbered by subscripts, i.e.,
 \begin{equation}\label{eq:15}
@@
 \begin{gather}\label{eq:18}
-B(q) = (t_s - t_{s-1})^{1/N} + \frac{(v_q - v_{q-1})^2} {m^2 (t_s - t_{s-1})^{1/N}} -2\frac{v_{q-1} + v_q}{m},\\
+B(q) = (t_s - t_{s-1})^{1/N} + \frac{(v_q - v_{q-1})^2} {m^2 (t_s - t_{s-1})^{1/N}} -2\frac{v_{q} + v_{q-1}}{m},\\
 \varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
@@
 \end{equation}
-i.e., completes the search when the length of the interval with maximal characteristic (\ref{eq:21}) becomes less than the predefined accuracy $\varepsilon > 0$.
+i.e., it completes the search when the length of the interval with maximal characteristic (\ref{eq:21}) becomes less than the predefined accuracy $\varepsilon > 0$.
 
@@
 In the course of optimization the algorithm builds the trial points $t^1,...,t^n$ which correspond to the points $x^1 = x(t^1),...,x^n=x(t^n)$ in the domain $Q$ .
-At these points the values $v^j = \Phi_\lambda(x^j), \; 1 \leq j \leq n,$ of the function $\Phi_\lambda(x)$ from (\ref{conv_problem}) are computed as values $\varphi(t^j)$ of the function $\varphi_t$  minimized by ALR.
+At these points the values $v^j = \Phi_\lambda(x^j), \; 1 \leq j \leq n,$ of the function $\Phi_\lambda(x)$ from (\ref{conv_problem}) are computed as values $\varphi(t^j)$ of the function $\varphi(t)$  minimized by ALR.
 
@@
 
-This information can be used for decreasing the number of trials during solving the problem (\ref{conv_problem}) with other weight vector $\lambda$.
+This information can be used for decreasing the number of trials during solving the problem (\ref{conv_problem}) with other weight vector.
 Indeed, if we take the weight vector $\mu = (\mu_1,...,\mu_k) \neq \lambda$, it is easy on the base of matrix $A$ to calculate values $\Phi_\mu(x^j), \; 1 \leq j \leq n,$ at points $x^1 = x(t^1),...,x^n=x(t^n)$, i.e.
@@
 Each group contained a series of MCO problems of the same dimension and number of criteria.
-Since all GKLS functions have a global minimum value equal to $-1$, the partial  criteria $w_i(x), \; 1 \leq i \leq k,$  increased by a positive constant to ensure the requirements (\ref{positive}).
+Since all GKLS functions have a global minimum value equal to $-1$, the partial  criteria $w_i(x), \; 1 \leq i \leq k,$  were increased by a positive constant to ensure the requirements (\ref{positive}).
 

### 2022-01-27 2022/Sensors/sensors.tex
Sensors paper template

--- 2021-09-25 2021/Entropy/entropy.tex
@@
 %=================================================================
-\documentclass[entropy,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+\documentclass[sensors,article,submit,moreauthors,pdftex]{Definitions/mdpi}
 
@@
 \articlenumber{0}
-\pubyear{2021}
-\copyrightyear{2020}
+\pubyear{2022}
+\copyrightyear{2022}
 %\externaleditor{Academic Editor: Firstname Lastname} % For journal Automation, please change Academic Editor to "Communicated by"
@@
 % Full title of the paper (Capitalized)
-\Title{Acceleration of Global Optimization Algorithm by Detecting Local Extrema Based on Machine Learning}
+\Title{Title}
 		   %Acceleration of Global Optimization Algorithm by Detecting Local Extrema Based on Machine Learning
@@
 \newcommand{\orcidauthorA}{0000-0001-5273-2471} % Add \orcidA{} behind the author's name
-\newcommand{\orcidauthorB}{0000-0002-8736-0652} % Add \orcidB{} behind the author's name
-\newcommand{\orcidauthorC}{0000-0001-6776-0096} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorB}{0000-0003-1809-7173} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorC}{0000-0003-1542-7624} % Add \orcidC{} behind the author's name
 
 % Authors, for the paper (add full first names)
-\Author{Konstantin Barkalov $^{1}$*\orcidA{} , Ilya Lebedev $^{1}$\orcidB{} and Evgeny Kozinov $^{1}$\orcidC{}}
+\Author{Konstantin Barkalov $^{1}$*\orcidA{} , Anton Shtanyuk $^{1}$\orcidB{} and Alexander Sysoyev $^{1}$\orcidC{}}
 
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Konstantin Barkalov, Ilya Lebedev and Evgeny Kozinov}
+\AuthorNames{Konstantin Barkalov, Anton Shtanyuk and Alexander Sysoyev}
 
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Barkalov, K.; Lebedev, I.; Kozinov, E.}
+\AuthorCitation{Barkalov, K.; Shtanyuk, A.; Sysoyev, A.}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
@@
 \address[1]{%
-$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhny Novgorod, Russia; ilya.lebedev@itmm.unn.ru (I.L.), evgeny.kozinov@itmm.unn.ru (E.K.)}
+$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhny Novgorod, Russia; anton.shtanyuk@itmm.unn.ru (A.Sh.), sysoyev@vmk.unn.ru (A.S.)}
 %$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
@@
 \abstract{
-The paper features the study of global optimization problems and numerical methods of their solution.
-Such problems are computationally expensive since the objective function can be multi-extremal, nondifferentiable, and, as a rule, given in the form of a ``black box''.
-The study used a deterministic algorithm for finding the global extremum.
-The said algorithm is based neither on the concept of multistart, nor nature-inspired algorithms.
-The article provides computational rules of the one-dimensional algorithm and the nested optimization scheme which could be applied for solving multidimensional problems.
-Note that the solution complexity of global optimization problems essentially depends on the presence of multiple local extrema.
-In this paper, we apply machine learning methods to identify regions of attraction of local minima.
-The use of local optimization algorithms in the selected regions can significantly accelerate the convergence of global search as it could reduce the number of search trials in the vicinity of local minima.
-The results of computational experiments carried out on several hundred global optimization problems of different dimensionalities presented in the paper confirm the effect of accelerated convergence (in terms of the number of search trials required to solve a problem with a given accuracy).
+abstract
 }
@@
 % Keywords
-\keyword{global optimization; local optimization; multiextremal problems; numerical methods; approximation; decision trees}
+\keyword{space-filling curves}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Diversity
-%\LSID{\url{http://}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Applied Sciences:
-%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Data:
-%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
-If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
-In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
-
-%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Toxins
-%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Encyclopedia
-%\encyclopediadef{Instead of the abstract}
-%\entrylink{The Link to this entry published on the encyclopedia platform.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
@@
 
-%The introduction should briefly place the study in a broad context and highlight why it is important.
-It should define the purpose of the work and its significance.
-The current state of the research field should be reviewed carefully and key publications cited.
-Please highlight controversial and diverging hypotheses when necessary.
-Finally, briefly mention the main aim of the work and highlight the principal conclusions.
-As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
-Citing a journal paper \cite{ref-journal}.
-Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}.
-Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Histories, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences.
-
-The successful application of machine learning (ML) methods to solve a wide range of problems leads to the emergence of new ways to apply ML for many tasks. 
-Methods of machine learning were shown to be particularly effective for identifying the principal properties of the phenomena (for example, physical, economic, or social), which are stochastic by nature or contain some hidden parameters \cite{Golovenkin2020,Gonoskov2019}.
-ML is also successfully used to solve complex problems of computational mathematics, for example, for simulation of dynamical systems  \cite{Seleznev2019}, solution of ordinary, partial, or stochastic differential equations \cite{Lagaris1998,Blechschmidt2021,Xu2020}.
-
-Particularly, ML could be applied for solving such a complex problem of computational mathematics as global optimization.
-The solution to this class of problems, as a rule, cannot be found analytically and, therefore, one needs to construct numerical methods to solve it.
-
-The numerical solution of optimization problems is fraught with significant difficulties.
-In many ways, they are related to the dimensionality and type of the objective function.
-Consequently, the most difficult problems are those in which the objective function is multi-extremal, nondifferentiable, and, moreover, given in the form of a ``black box'' (i.e., in the form of some computational procedure, the input of which is an argument, and the output is the corresponding value of the function).
-These complex problems are the main focus of this article.
-
-There are several approaches to the construction of numerical methods for solving global optimization problems. 
-Some algorithms are based on the idea of a multistart: launching a local search either from different starting points or with varying parameters.
-Local optimization methods have a high convergence rate.
-At the same time, one of the main problems in multistart schemes is the choice of starting points that would correspond to the regions of attraction of various local solutions.
-Machine learning methods can be successfully applied to solve this problem.  
-For example, in \cite{RinnooyKan1987} methods of cluster analysis were used to select promising starting points.
-In \cite{Cassioli2012} the area for starting the local method was allocated based on the classification of starting points using a support vector machine.
-
-%Russian
-{ \color{red}
-Machine learning methods are actively used in combination with bayesian optimization algorithms based on the probabilistic surrogate models of the objective function.
-The detailed overview of this trend in the development of global optimization methods is presented in \cite{Archetti2019,Zhigljavsky2021}.
-}
-
-Another popular class of methods for solving global optimization problems is metaheuristic algorithms. 
-Many of them are based on imitation of the processes occurring in living nature.
-The parameters of such algorithms could also be tuned using ML.
-For example, \cite{Jin2005} provides an overview of machine learning applications in evolutionary algorithms.
-
-Note that the algorithms of the latter class do not provide guaranteed convergence to the solution of the problem and are inferior to deterministic algorithms in terms of the quality of solution \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a particular set).
-Therefore, deterministic methods seem to be potentially more effective.
-
-%Russian
-This paper aims to further develop the efficient deterministic global optimization method known as \textit{the information-statistical global search algorithm} \cite{Strongin2000}.
-The book referenced here contains the results of theoretical studies of the method that are of direct importance for its practical implementation.
-In particular, it discusses in detail the issues of convergence, the choice of parameters, and the conditions for stopping the algorithm, etc.
-Note that the global search algorithm was originally designed for solving unconstrained optimization problems.
-Later, it was generalized to solve problems with non-convex constraints\cite{Barkalov2002} and multicriteria optimization problems \cite{Gergel2020}.
-At the same time, scholars proposed various parallel versions of these algorithms, which can be used on modern supercomputers \cite{Barkalov2016,globalizerSystem,Strongin2018}.
-
-Several strategies have been proposed to speedup the global search algorithm (in terms of the number of iterations required to solve the problem with a given accuracy).
-In this paper, we propose a new approach to acceleration based on identifying areas of attraction of local minima using machine learning methods.
-The identification of regions of attraction and the launch of local search in these regions can significantly reduce the number of trials required for the method to achieve global convergence.
-Experiments carried out on a series of several hundred test problems confirm this statement.
+\cite{Strongin2000}
+Introduction
 
 \section{Problem Statement}
-
-In this paper, we will consider global optimization problems of the form
-\begin{eqnarray}\label{main_problem}
-& \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
-& D=\left\{y\in \text{R}^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
-\end{eqnarray}
-Problem (\ref{main_problem}) is considered under the assumption that the objective function is multi-extremal, is given in a form of a ``black box'', and the calculation of its values is associated with solving the problem of numerical simulation, which makes the solution a labor-intensive operation.
-
-A typical situation for many applied problems is when a limited change in the vector of parameters $y$ causes a limited change in the values of $\varphi(y)$.
-The mathematical model describing this premise is based on the assumption that the Lipschitz condition is satisfied
-\[
-\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in D,\; 0<L<\infty.
-\]
-This assumption is typical for many approaches to the development of optimization algorithms \cite{Jones1993,Pinter1996,Zilinskas2008,Evtushenko2009,Sergeyev2020}.
-At the same time, many known approaches are based on various methods of dividing the search domain into a system of subdomains and then choosing the most promising subdomain for placing the next trial (calculating the value of the objective function)  \cite{Jones2009,Zilinskas2010,Evtushenko2013,Kvasov2013,Paulavicius2016,Paulavicius2020}.
-An important property of global optimization problems is the fact that, in contrast to the problems of finding a local extremum, the global minimum is an integral characteristic of the problem being solved.
-Making sure that the point  $y^*\in D$ is a solution to the problem requires going beyond its neighborhood to the investigation of the entire search domain.
-As a result, when minimizing substantially multi-extremal functions, the numerical method must construct a coverage of the search domain.
-The number of nodes of this coverage increases exponentially with increasing dimensionality.
-This feature determines the high complexity of solving multiextremal optimization problems making dimensionality a critical factor that affects the complexity of their solving.
-
-The dimensionality in multi-extremal optimization leads to many issues, so, scholars use a wide variety of approaches to reducing it.  
-For example, simplicial or diagonal partition of the search domain allows using methods for solving one-dimensional problems to solve the original multidimensional problem (see, for example,  \cite{PaulaviciusZilinskas2014,Sergeyev2017}).
-Another well-known approach to dimensionality reduction is using the Peano space-filling curves to map the multidimensional domain onto a one-dimensional interval \cite{Strongin2000,Sergeyev2013}.
-
-In this work, we will use another method based on the nested optimization scheme \cite{Shi2000,Grishagin2001,VanDam2010,Grishagin2015} and its generalization \cite{Grishagin2016,Grishagin2016_1}.
-The nested optimization scheme, on the one hand, does not worsen the properties of the objective function (unlike reduction using Peano curves), and, on the other hand, does not require the use of complex data structures to support simplex or diagonal partitions of the feasible region.
-At the same time, the nested optimization scheme makes it possible to reduce the original multidimensional optimization problem to a family of recursively connected one-dimensional optimization subproblems, which can be solved by a wide range of one-dimensional global optimization algorithms.
 
@@
 
-As a standard, let us consider a one-dimensional multiextremal optimization problem
-\begin{equation}\label{uni_problem}
-\varphi^\ast = \varphi(x^\ast)=\min{\left\{\varphi(x):x\in \left[a,b\right], a<b, a,b \in R
-\right\}}
-\end{equation}
-with an objective function satisfying the Lipschitz condition.
-
-Here is the description of global search algorithm (GSA) for solving the basic problem in accordance with \cite{Strongin2000}.
-In the course of its work, GSA generates a sequence of points  $x^i$, at which the values of the objective function  $z^i=\varphi(x^i)$ are calculated. 
-We will refer to the process of calculating the value of the objective function as \textit{trial}.
-
-In accordance with the algorithm, the first two trials are carried out at the boundary points of the segment $[a,b]$, i.e. $x^0=a,\;x^1=b$. 
-At these points, the values of the objective function  $z^0=\varphi(x^0),\;z^1=\varphi(x^1)$ are calculated and the counter value is set to $k=1$. 
-The point of the next trial $x^{k+1}, k\geq 1,$ is selected in accordance with the following procedure.
-
- Step 1. Renumber  (starting at 0) the points $x^i,\:0\leq i\leq k$, of the trials conducted  in ascending order of the coordinate, i.e.
-\begin{equation}\label{xt}
-a=x_0<x_1<\ldots <x_{k}=b.
-\end{equation} 
-Associate the values of the objective function $z_i=\varphi(x_i), \; 0\leq i\leq k,$ to the points $x_i, \; 0\leq i\leq k$, at which these values were calculated.
-
-Step 2. Calculate the maximum absolute value of the relative first difference
-\begin{equation}\label{mu}
-\mu=\max_{1\leq i\leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
-\end{equation}
-where $\Delta_i = x_i-x_{i-1}$. If the value calculated in accordance with (\ref{mu}) is equal to zero, then take $\mu = 1$.
-
-Step 3. For all the intervals $(x_{i-1},x_i),1\leq i\leq k$,  calculate the value
-\begin{equation}\label{R}
-R(i)=r\mu\Delta_i+\frac{(z_i-z_{i-1})^2}{r\mu\Delta_i}-2(z_i+z_{i-1}),
-\end{equation} 
-refered to as the \textit{characteristics} of the interval; value $r>1$ is the parameter of the algorithm.
-
-Step 4. Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
-\begin{equation}\label{MaxR}
-R(t)=\max_{1\leq i\leq {k}}R(i).
-\end{equation}
-If the maximum characteristic corresponds to several intervals, then choose the minimum number that satisfies (\ref{MaxR}) as $t$.
-
-Step 5. Carry out a new trial at the point
-\begin{equation}\label{xk1}
-x^{k+1}=\frac{1}{2}(x_{t-1}+x_t) - \frac{z_t-z_{t-1}}{2r\mu}.
-\end{equation}
-
-The algorithm stops when the condition  $\Delta_t<\epsilon$ is satisfied; here $t$ is from (\ref{MaxR}), and $\epsilon>0$ is a given accuracy. 
-The values
-\[
-z_k^\ast=\min_{0\leq i \leq k}\varphi(x^i), \ x_k^\ast=\arg \min_{0\leq i \leq
- k}\varphi(x^i)
-\] 
-are selected to estimate the solution
-
-The theoretical conditions that determine the convergence of the algorithm are presented in \cite{Strongin2000}. 
-The work of the algorithm during the minimization of a specific multiextremal function, which is specified in accordance with formula  (\ref{hill}), is shown in Figure \ref{fig1}.
-The algorithm was launched with the parameter $r=2.2$ from (\ref{R}) and value $\epsilon = 10^{-3}$ in the method stopping condition.
-Figure \ref{fig1} shows the objective function graph and points of 71 search trials which GSA needed to solve the problem to the specified accuracy.
-This highlights the problem of all methods of global optimization --- the concentration of trial points in the vicinity of local minima of the problem, which are not a global solution.
-
-\begin{figure}[H]
-\includegraphics[width=1.0\linewidth]{HillAGP90.png}
-\caption{Minimization of a function of the form  (\ref{hill}) using GSA.}
-\label{fig1}
-\end{figure}
-
-\subsection{Machine Learning Regression as a tool for identifying attraction regions of local extrema}\label{TreeGSA}
-
-The functions considered in this study belong to the class of Lipschitzian functions.
-Therefore, classical regression methods (for example, polynomial regression, where a function is approximated by a polynomial of a given degree) will not properly match the behavior of the function.
-A more powerful tool for this task is regression splines.
-When constructing a regression spline, the domain is divided into $K$ non-overlapping subdomains.
-In each of such domains, the function is approximated by a polynomial.
-Dividing the interval into a sufficient number of subdomains allows one to very accurately approximate the original function.
-
-Regression can also be constructed using such a powerful tool as artificial neural networks. 
-Different types of networks can be used to build the regression, for example, multilayer perceptron, radial basis function network, etc.
-However, in both of these cases, the model itself (spline or neural network) becomes rather complex for the analysis required to solve the given problem (identifying areas of attraction of local extrema).
-
-Therefore, within the framework of the study, we chose a regression model based on decision trees to analyze the local behavior of a function.
-For example, if the objective function is properly approximated by polynomials, then the polynomial regression, of course, will appropriately convey the properties of the function.
-However, if there is a more complex relationship between them, then the decision tree can surpass the classical variants of regression in terms of the quality of the approximation.
-At the same time, the regression based on the decision tree makes it possible to easily identify the areas of attraction of local extrema with sufficient accuracy.
-
-Building a regression using a decision tree consists of two main steps:
-\begin{itemize}
-	\item Search domain $D$ is divided into $J$ non-overlapping subdomains  $D_1, D_2, ..., D_J$, provided that $D = \bigcup_{j=1}^{J}{D_j}$.
-	\item Any value falling into the subdomain $D_j$, i.e. $x \in D_j$, is matched to the average value  $c_j$ based on the training trials that fall into this subdomain. 
-\end{itemize}
-In fact, decision trees build a model of a function of the form
-\begin{equation}\label{dtree}
-f(x) = c_j, \; x\in D_j.
-\end{equation}
-
-%Russian
-{\color{red}
-Generally speaking, a decision tree is a binary tree, the leave nodes of which contain the values of the function, and the other nodes contain the transition conditions.
-In our case, when applying a decision tree to construct a regression, each node corresponds to the results of several trials and the value $c_j$, which is calculated as the mean square of the trial points assigned to this node.
-The final piecewise constant approximation is constructed using the values $c_j$, located in the leave nodes of the tree.
-
-The tree is built recursively, starting from the root node. 
-A decision rule is applied at each node: all data is divided into two groups (according to a partitioning rule) and sent to the left and right child nodes.
-The procedure then recursively separates the left and right nodes.
-Recursion stops at a node in one of the following cases:
-\begin{itemize}
-	\item The number of trial points assigned to the node becomes less than the specified threshold value (we used 1).
-	\item The sum of the squared deviations of the function values from the value $c_j$, assigned to this node becomes less than the set accuracy (we used $10^{-3}$).
-\end{itemize}
-For more information on the algorithm for constructing regression using decision trees, see, for example \cite{Breiman1984}.
-}
-
-Regression built using decision trees,
-is, on the one hand, rather simple, and on the other hand, it adequately reflects the properties of the function under investigation (the presence or absence of local minima).
-The selection of regions of attraction of local minima using model (\ref{dtree}) can be organized as follows.
-
-Let $x^{k+1}$ be the point of the current trial. For a given point, an index  $j$ is sought such that $x^{k+1} \in D_j$. 
-Next, the  $c_j$ values corresponding to neighboring subdomains are compared. If one of the conditions is met
-\begin{gather}
-c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3} \leq c_{j+4}, \; j=1; \nonumber \\ 
-c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2} \leq c_{j+3}, \; j=2;  \nonumber \\  
-c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1} \leq c_{j+2}, \; 3 \leq j \leq J-2; \label{LPC} \\ 
-c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j} \leq c_{j+1}, \; j=J-1; \nonumber \\ 
-c_{j-4} \geq c_{j-3} \geq c_{j-2} \geq c_{j-1} \geq c_{j}, \; j=J; \nonumber 
-\end{gather}
-then the subdomain $D_j$ is considered the area of attraction of the local minimum. Here, one can start a local search, and subsequently exclude this subdomain from the global search. 
-Any of the zero-order local methods can be used as a local algorithm, for example, golden section search or parabolic interpolation \cite{Press}.
-
-In order to modify the global search algorithm from subsection  \ref{CoreGSA} to exclude the regions of attraction of local minima, we associate each trial point  $x^i$ obtained during the operation of the algorithm with an additional attribute $q^i \in \{0,1,2\}$, which will characterize the properties of this point.
-The value  $q^i=0$ is assigned by default and indicates that point $x^i$ is obtained as a result of rule (\ref{xk1}) of the global search algorithm.
-The value $q^i=1$ is assigned to the points obtained as a result of the work of the local method, while the value  $q^i=2$ corresponds to the point of the local minimum found as a result of the work of the local search.
-
-%Russian
-{\color{red}
-Using the additional attribute $q^i, \; 0 \leq i \leq k ,$ makes it possible to distinguish between points obtained during a global ($q^i=0$) and a local ($q^i=1,2$) search.
-The intervals, the boundary points of which have the value  $q^i=1$ or $q^i=2$, will be skipped in the global search, i.e.
-they will not be tried further.
-While the selected value $q^i=2$, which corresponds to the local minimum will allow using these points for checking the stopping criterion.
-When the rules of the global search signal that a new trial point needs be placed in a given  $\epsilon$-neighborhood of one of the found local minima, this will correspond to the end of the search.
-}
-
-Let us now describe a modified global search algorithm that uses decision trees to isolate and exclude areas of attraction of local minima; we will further refer to this algorithm as GSA-DT.
-Recall that the superscript corresponds to the number of the iteration at which the trial was carried out at this point, and the subscript corresponds to the number of the point in the row  (\ref{xt}).
-
-Steps 1-5 of the GSA-DT algorithm are the same as steps 1-5 of GSA.
-
-Step 6. If $q_{t-1} \in \{1,2\}$ or $q_t \in \{1,2\}$, then go to check the stopping criterion. Otherwise, go to Step 7.
-
-Step 7.
-Construct a decision tree based on the results of the trials performed, and obtain the corresponding piecewise constant approximation, which assigns the value $c_1, c_2, ..., c_J$ to each subdomain  $D_1, D_2, ..., D_J$.
-
-Step 8.
-For the point $x^{k+1}$ of the current trial, find a number $j$ such that $x^{k+1} \in D_j$ and check whether the condition  (\ref{LPC}) is satisfied.
-If condition (\ref{LPC}) is satisfied, start a local search in the domain $D_j$ from the point  $x^{k+1}$.
-The results of all trials performed during the local search at the points $x^{k+2}, ...,x^{k+k_{local}}$ are stored in the information base of the algorithm and are used at subsequent iterations. 
-All these points receive the attribute  $q^i=1$, $i = k+2, ... , k+k_{local}$. The attribute equal to 2 is assigned to the point corresponding to the found local minimum.
-
-The stopping criterion of the modified algorithm will look as follows. 
-The algorithm stops when one of the following conditions is met:
-
-\begin{itemize}
-	\item $|x_{t} - x_{t-1}|<\epsilon$;
-	\item $|x^{k+1} - x_{t-1}|<\epsilon$ and $q_{t-1} = 2$;
-	\item $|x^{k+1}-x_{t}|<\epsilon$ and $q_{t} = 2$;
-\end{itemize}
-where $t$ is from (\ref{MaxR}), and $\epsilon>0$ is a given accuracy.
-
-%Russian
-{\color{red}
-Note that the specified stopping criterion is checked after Step 6, i.e.
-after the calculation according to the rules of the global search for the point of the next trial $x^{k+1}$, but before the trial itself is carried out in it.
-The idea behind this criterion is as follows.
-The search stops when the interval  $(x_{t-1},x_t)$, in which the point $x^{k+1}$ falls, becomes sufficiently small, or when the point $x^{k+1}$ falls into a small neighborhood of one of the found local minima.
-Such minima in this case will be simultaneously global.
-}
-
-For example, consider the work of the GSA-DT algorithm with minimization of the same multi-extremal function, which is presented in Figure \ref{fig1}.
-The same parameters were used at the start of the algorithm: the parameter  $r=2.2$ from (\ref{R}) and value $\epsilon = 10^{-3}$ in the stopping criterion of the method.
-Figure \ref{fig2} illustrates the operation of the GSA-DT algorithm.
-In addition to the graph of the objective function, it shows a piecewise constant approximation of the form (\ref{dtree}) built at the final stage of the search.
-Black points on the graph correspond to the global search phase, green points correspond to the work of the local method.
-In total, the GSA-DT method required 49 trials to solve the problem; there was no accumulation of trial points in the vicinity of local minima.
-
-\begin{figure}[H]
-\includegraphics[width=1.0\linewidth]{HillTree90.png}
-\caption{Using GSA-DT to minimize the function (\ref{hill}) }
-\label{fig2}
-\end{figure}
-
-\subsection{Adaptive Dimension Reduction Scheme}
-
-The recursive nested optimization scheme is based on the well-known relation  \cite{Grishagin2001} 
-\begin{equation}\label{nested}
-\min_{y \in D}\varphi(y) = \min_{y_1\in\left[a_1,b_1\right]}\min_{y_2\in\left[a_2,b_2\right]}...\min_{y_N\in\left[a_N,b_N\right]}\varphi(y),
-\end{equation}
-which allows reducing the solution of the original multidimensional problem (\ref{main_problem}) to the solution of a family of recursively connected one-dimensional subproblems.
-
-For a formal description of the nested optimization scheme, we introduce a family of functions defined in accordance with the relations
-\begin{equation}\label{nested_N}
-\varphi_N(y_1,...,y_N) \equiv \varphi(y_1,...,y_N),
-\end{equation}
-\begin{equation}\label{nested_i}
-\varphi_i(y_1,...,y_i) = \min_{ y_{i+1} \in\left[a_{i+1},b_{i+1}\right]} \varphi_{i+1}(y_1,...,y_i,y_{i+1}), 1\leq i\leq N-1.
-\end{equation}
-
-Then, in accordance with  (\ref{nested}), solving the multidimensional problem  (\ref{main_problem}) is reduced to solving a one-dimensional problem 
-\begin{equation}\label{nested_1}
-\varphi^* = \min_{y_1\in\left[a_1,b_1\right]}\varphi_1(y_1).
-\end{equation}
-However, each calculation of the value of the function $\varphi_1$ at some fixed point $y_1$ presupposes the solution of the one-dimensional optimization problem of the second level
-\begin{equation}
-\varphi_1(y_1) = \min_{y_2\in\left[a_2,b_2\right]}\varphi_2(y_1,y_2).
-\end{equation}
-Calculation of the values of the function  $\varphi_2$ in turn, requires one-dimensional minimization of the function  $\varphi_3$ all the way to the solution of the problem
-\begin{equation}
-\varphi_{N-1}(y_1,...,y_{N-1}) = \min_{ y_{N} \in\left[a_{N},b_{N}\right]} \varphi_{N}(y_1,...,y_{N})
-\end{equation}
-at the last level of recursion.
-
-The solution of the set of subproblems arising in the nested optimization scheme  (\ref{nested_i}) can be organized in different ways. 
-The obvious way (described in detail in \cite{Grishagin2001,Grishagin2015}) is based on solving subproblems in accordance with the recursive order of their generation.
-However, a significant part of the information about the objective function is lost here.
-
-Another approach is an adaptive scheme, in which all subtasks are solved simultaneously, which allows taking into account much more information about a multidimensional problem, thereby speeding up the process of its solution.
-This approach was theoretically substantiated and tested in  \cite{Grishagin2016,Grishagin2016_1,Grishagin2018}.
-
-Note that within the framework of the original nested optimization scheme, the generated subproblems are solved only sequentially; the resulting hierarchical scheme for generating and solving subproblems has the form of a tree.
-The construction of this tree occurs dynamically in the process of solving the original problem (\ref{main_problem}).
-In this case, the calculation of one value of the function $\varphi_i(y_1,y_2,...,y_i)$ at the $i$-th level requires a complete solution of all problems of one of the subtrees of level  $i+1$.
-
-The adaptive nested optimization scheme of dimensionality reduction changes the order of solving subproblems: they will be solved not one by one (in accordance with their hierarchy in the problem tree), but simultaneously, i.e.
-there will be a set of subtasks in the process of solution.
-Within the adaptive scheme:
-\begin{itemize}
-	\item 
-to calculate the value of $i$-th level function from (\ref{nested_i}) a new $i+1$ level problem is generated, in which only one trial is carried out, after which the new generated problem is included in the set of already existing problems to be solved;
-	\item 
-	iteration of the global search consists in choosing one (most promising) problem from the set of available problems, in which one trial is carried out; the new trial point is determined according to the basic global search algorithm from subsection \ref{CoreGSA} or a modified algorithm from subsection  \ref{TreeGSA};
-	\item
-the minimum values of functions from  (\ref{nested_i}) are their current estimates obtained based on accumulated search information.
-\end{itemize}
-
-%A brief description of the main steps of the adaptive dimensionality reduction scheme is as follows.
-%Let nested subproblems of the form (\ref{nested_i}) be solved using the global search algorithm described in subsection  \ref{CoreGSA}.
-Then each subproblem  (\ref{nested_i})  can be assigned some numerical value called a characteristic of this problem.
-As such a characteristic, we can take the value $R(t)$ from (\ref{MaxR}), i.e.
-the maximum characteristic from the characteristics of the intervals formed in this task.
-The higher the values of this characteristic, the more promising is the search for the global minimum of the original problem (\ref{main_problem}) in the given subproblem.Therefore, at each iteration, a subproblem with the maximum characteristic is selected, and a trial is carried out in it.
-This either leads to the calculation of the value of the objective function  $\varphi(y)$ (if the selected subtask belonged to the level  $j=M$), or generates new subproblems according to (\ref{nested_i}) for $j\leq M-1$.
-In the latter case, new generated tasks are added to the current set of tasks, their characteristics are calculated, and the process is repeated.
-The completion of the optimization process occurs when the condition for stopping the algorithm that solves this problem is satisfied for the root problem.
-
 \section{Experimental Results}
-
-Numerical experiments were performed on the Lobachevsky supercomputer of the University of Nizhny Novgorod (operating system CentOS 7.2, management system SLURM).
-One supercomputer node has 2 Intel Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM.
-The CPU is 8-core (i.e.
-a total of 16 CPU cores are available on the node).
-All the algorithms were implemented in C++; GCC 5.5.0 was used for compilation on the supercomputer.
-
-The traditional approach to assessing the effectiveness of global optimization methods is based on using these methods to find the numerical solution of a series of problems. 
-In this case, the assumption is that a certain algorithm is used to generate the next problem to be solved. 
-Typical examples of such test function classes are Shekel and Hill functions. 
-The first of them (denoted $F_{SH}$) is based on the formula
-\begin{equation}\label{shekel}
-  \varphi(x)=-\sum_{j=1}^{10}\frac{1}{(K_j(x-A_j)^2+C_j)},\;  x\in[0,10],
-\end{equation}
-where parameters $1\le K_j\le 3,\: 0 < A_j,\: C_j < 10, \;$ are independent random variables uniformly distributed in the indicated intervals. 
-The next generator (denoted   $F_{HL}$) is determined by the expression
-\begin{equation}\label{hill}
-  \varphi(x)=\sum_{j=1}^{14}(A_j\sin(2j\pi x) + B_j\cos(2j\pi x)),\: x\in[0,1],
-\end{equation}
-where the values of the parameters $A_j,\: B_j,\: 1 \le j \le 14$, are independently and uniformly distributed in the interval  $[-1,1]$.
-
-Let us compare the basic global search algorithm (GSA) and its decision tree-based modification (GSA-DT) with the well-known DIRECT global optimization algorithm  \cite{Jones2009}.
-{\color{red}
-%Russian 
-The choice of this particular method for comparison is explained as follows.
-DIRECT is one of the most well-known and popular deterministic methods for solving global optimization problems with a `` black box '' objective function.
-An overview of various modifications of the method, as well as examples of solving problems, is given in \cite{Jones2021}.
-It is known that with a sufficiently large number of search trials, DIRECT is guaranteed to find a global solution to the problem. 
-However, if the stopping by accuracy is used as a criterion, then the method can abruptly stop at one of the local minima, which is confirmed by the experimental results presented in this section.
-
-Regarding the deterministic optimization algorithms implemented in popular Computer Algebra Systems (CAS), their comparison with methods such as DIRECT or GSA will be incorrect for the following reason.
-Optimization algorithms from CAS are focused on solving problems with an objective function set explicitly, in the form of a formula.
-Formula definition of the objective function assumes that its derivatives are also known, which makes it possible to use first-order methods with significantly faster convergence than zero-order methods, which include both DIRECT and GSA.
-However, first-order methods do not guarantee finding the global minimum.
-For example, the Mathematica system offers methods for solving global optimization problems that are guaranteed to find a global solution to the problem only if the objective function and constraints are linear or convex.
-Otherwise, the result may sometimes only be a local minimum.
-
-Moreover, methods that require a formulaic specification of the objective function cannot be applied to the solution of a large class of applied problems in which the form of the objective function is not known, and its values are calculated as a result of numerical simulation.
-The use of heuristic methods for solving problems of this kind (such as Differential Evolution, Simulated Annealing or Random Search) also does not always lead to the solution of the problem.
-These methods often do find the global minimum, but are not guaranteed to do so.
-In terms of the number of correctly solved problems, heuristic methods are inferior to deterministic ones \cite{Sergeyev2018}.
-}
-
-The decision trees in the GSA-DT algorithm were built using the OpenCV 4.5.1 library (class cv::ml::DTrees).
-A regression function was constructed by a single decision tree.
-This allowed us to obtain a piecewise constant approximation of the objective function in which one or several trial points corresponded to each leaf node of the tree.
-The tree was built without any limitations on maximum depth (MaxDepth); the accuracy of tree construction (RegressionAccuracy) was $10^{-3}$; the minimum number of trial points in the tree node (MinSampleCount) was equal to one; all other parameters were set by default.
-Stopping tree building occurred if all absolute differences between an estimated value in a node and values of train samples in this node are less than accuracy.
-
-The global optimization methods discussed above were compared when solving 100 problems from the  $F_{SH}$ and $F_{HL}$ classes.
-The problem will be considered correctly solved if after stopping the method by accuracy (i.e.
-when the length of the current search interval becomes less than $\epsilon \cdot \left|b-a\right| $) the current estimate of the optimum  $x_k^*$ lies in the $\epsilon$-neighborhood of the known solution of the problem  $x^*$, i.e.
-if the condition  $|x^*-x_k^*| \leq \epsilon \cdot \left|b-a\right|$ is satisfied.
-
-Tables \ref{table:average_Shekel} and \ref{table:average_Hill} show the number of search trials that, on average, were required to minimize the Shekel and Hill functions with different search accuracy  $\epsilon$.
-The number of unsolved problems is indicated in parentheses.
-{\color{red}
-These and subsequent tables feature the total number of trials that were performed during both global and local searches by GSA-DT.
-When solving problems with an accuracy  $10^{-2}$, $10^{-3}$, $10^{-4}$ the number of local search launches when solving one problem on average was equal to 2, 3, and 4, respectively; and the ratio of the number of trials performed according to the global search rules to the number of trials performed by the local method was (with appropriate accuracy) 7.2, 6.0, 4.5 for problems of the $F_{HL}$ class and 3.2, 2.3, 1.9 for the $F_{SH}$ class problems.
-}
-
-%Ebglish
-The experimental results show that with a rough solution to the problem, all methods show similar results in terms of the number of trials, while with a high solution accuracy, the GSA-DT algorithm requires 2 times fewer trials than its prototype.
-At the same time, GSA-DT outperforms the DIRECT method both in the average number of search trials and in the number of correctly solved problems.
-In particular, if we use the accuracy $\epsilon = 10^{-2}\left|b-a\right|$, then the DIRECT method stops too early and does not find a global solution to many problems.
-Therefore, in further experiments in which multidimensional problems are solved, we will not use DIRECT for comparison, since when solving multidimensional problems with stopping by accuracy, this method will provide a correct solution to no more than  $50\%$ of problems.
 
@@
 
-\begin{specialtable}[H] 
-	\caption{The average number of tests when minimizing Hill test functions (the number of unsolved problems is indicated in parentheses))}\label{table:average_Hill}
-	\center
-\begin{tabular}{cccc}
-\toprule
-        & \textbf{$\epsilon = 10^{-4}$} & \textbf{$\epsilon = 10^{-3}$} & \textbf{$\epsilon = 10^{-2}$} \\
-\midrule					  
-DIRECT                & 66(12) & 36(31)  & 20(51)  \\
-GSA                   & 130    & 75      & 43      \\
-GSA-DT                & 64     & 59      & 50      \\
-\bottomrule
-\end{tabular}
-\end{specialtable}
-
-The next series of experiments involved the solution of multidimensional problems. 
-A well-known generator of multi-extremal optimization test problems is GLKS  \cite{Gaviano2003}.
-It can be used to generate test functions with given properties: the number of local extrema, their areas of attraction, the global minimum point, and the value of the objective function at this point, etc.
-The procedure for generating test functions is based on using polynomials to redefine a convex quadratic function (paraboloid).
-Test functions are defined by five parameters:
-\begin{itemize}
-	\item dimensionality of the problem  $N$;
-	\item the number of local minima $l$;
-	\item value of the global minimum $f^*$;
-	\item radius of the area of attraction of the global optimizer $\rho^*$;
-	\item the distance between the global optimizer and the vertex of the paraboloid $d^*$.
-\end{itemize}
-By changing the specified parameters, one can create test classes with different properties.
-For example, with a fixed dimensionality of the problem and the number of local minima, a more complex class can be generated by narrowing the region of attraction of the point of the global minimum or by increasing the distance between this point and the vertex of the paraboloid.
-In the experiments, the values $l=10$, $f^*=-1$, $\rho^*=0.2$ and $d^*=0.9$ were used.
-
-As an example, consider the operation of the GSA and GSA-DT algorithms when solving one of the two-dimensional problems generated by the GKLS generator. 
-The level lines of the objective function shown in Figures \ref{fig3} and \ref{fig4}, indicate the presence of ten local extrema. 
-When starting the algorithms, the same parameters were used:  $r=3.0$ from (\ref{R}) and $\epsilon = 10^{-2}\left|b-a\right|$ in the stopping criterion of the method. 
-Black dots in Figures \ref{fig3} and \ref{fig4} show the points of the search trials performed by the methods in the process of solving the problem.
-In this case, the GSA algorithm required 247 trials, while the GSA-DT algorithm took 138 trials.
-The red dot in the figures marks the exact solution of the problem, and the yellow dot indicates the best approximation found by the algorithm. 
-Green dots in Figure \ref{fig4} indicate trials performed as part of a local search. 
-This graph demonstrates that using decision trees to identify areas of attraction of local minima removes the problem of accumulation of test points in the region of local extrema inherent in the original global search algorithm.
-
-\begin{figure}[H]
-\includegraphics[width=0.6\linewidth]{GKLSAdaptiv6_30line.png}
-\caption{Using GSA algorithm to solve GKLS problems}
-\label{fig3}
-\end{figure}
-
-\begin{figure}[H]
-\includegraphics[width=0.6\linewidth]{GKLSTree6_30line.png}
+\begin{figure}
+\includegraphics[width=0.6\linewidth]{Definitions/logo-mdpi}
 \caption{Using GSA-DT algorithm to solve GKLS problems}
@@
 
-We used GKLS to generate 300 test problems of dimensionalities $N=2,3,4$ (100 problems of each dimensionality).
-The resulting series of problems were solved using the GSA and GSA-DT algorithms with the parameter $r=5.0$ from (\ref{R}).
-The specified value of the parameter $r$ ensures the solution of  $100\%$ of the problems; at lower values of the parameter, some problems were not solved correctly.
+some text
 
-Tables \ref{table:average_GKLS01} and \ref{table:average_GKLS002} show the average number of trials required by the GSA and GSA-DT methods to correctly solve all problems with an accuracy of $\epsilon = 10^{-2}\left|b-a\right|$ and $\epsilon = 2 \cdot 10^{-3}\left|b-a\right|$ respectively.
+ Example of a figure that spans the whole page width. The same concept works for tables, too.
 
-The data from the tables confirm that the global search algorithm based on the application of machine learning to extract local extremum provides a faster solution to multiextremum problems than the basic global search algorithm.
-For rough accuracy solutions, the acceleration is about  $30\%$, for high accuracy solutions, the process is accelerated from 2 to 6 times.
-
-\begin{specialtable}[H] 
-	\caption{Solving GKLS problems with an accuracy  $\epsilon = 10^{-2}\left|b-a\right|$}\label{table:average_GKLS01}
-	\center
-\begin{tabular}{cccc}
-\toprule
-        & $N=2$ & $N=3$  & $N=4$    \\
-\midrule
-GSA     &  937  &  12716 &  206869  \\
-GSA-DT  &  653  &  9204  &  156190  \\
-\bottomrule
-\end{tabular}
-\end{specialtable}
-
-\begin{specialtable}[H] 
-	\caption{Solving GKLS problems with an accuracy $\epsilon = 2 \cdot 10^{-3}\left|b-a\right|$}\label{table:average_GKLS002}
-	\center
-\begin{tabular}{cccc}
-\toprule
-
-       & $N=2$ & $N=3$  & $N=4$    \\
-\midrule
-GSA    & 1489  & 69764  & 583903  \\
-GSA-DT & 831   & 10776  & 173155   \\
-\bottomrule
-\end{tabular}
-\end{specialtable}
+\end{paracol}
+\begin{figure}	
+\widefigure
+\includegraphics[width=15 cm]{Definitions/logo-mdpi}
+\caption{This is a wide figure.\label{fig2}}
+\end{figure}  
+\begin{paracol}{2}
+\linenumbers
+\switchcolumn
 
@@
 \section{Conclusions and Future Work}
-
-The article discusses an efficient deterministic method for solving multiextremal optimization problems --- the information-statistical global search algorithm.
-A new way of speeding up the operation of this algorithm was proposed (in terms of the number of trials required to solve the problem with a given accuracy).
-This method is based on identifying areas of attraction of local minima of the objective function using machine learning methods.
-The identification of regions of attraction and the launch of local search in these regions can significantly reduce the number of trials required for the method to achieve global convergence.
-Within the framework of the investigated approach, solving multidimensional problems is reduced to solving a series of information-related one-dimensional subproblems; therefore, the key point is to identify local minima in one-dimensional problems.
-This is achieved by approximation of the objective function built using decision trees.
-Computational experiments were carried out on a series of test problems of different dimensions to compare the speed of the original global search algorithm (GSA) and its modification, which uses decision trees to identify local minima of the objective function (GSA-DT).
-The experimental results show that the use of the GSA-DT algorithm can significantly (up to 6 times) reduce the number of trials required to solve the problem with a given accuracy.
-
-Further research into this issue will focus on using more complex models of the objective function to obtain a more accurate approximation.
-We plan to use artificial neural networks as such an approximator.
-This will require the development of new methods for identifying local extrema since function approximation using a neural network is more complicated from this point of view.
-We also plan to pay attention to the issues of the reliability of the results obtained using machine learning methods.
-For the solved model problems, the use of machine learning methods shows good results, but the question of whether this effect will persist in more complex problems remains open.
-
-%Russian
-{\color{red}
-Another direction for further work will be the combination of the proposed approach to the application of local search with traditional methods of accelerating global optimization algorithms.
-For example, a well-known way to speed up algorithms of this class is to use local adaptive estimates of the Lipschitz constant $L_i$ in various subdomains $D_i \in D$ within the search domain $D$ instead of a single global estimate of the constant $L$ for the entire domain $D$ ( see \cite{Kvasov2020,Sergeyev2021}).
-This allows the algorithm to adjust to the local behavior of the objective function in different parts of the feasible domain, and, thereby, reduces the number of search trials required to achieve convergence.
-Using only global information about the behavior of the objective function during its minimization can significantly slow down the convergence of the algorithm to a point of global minimum.
-Further research on this issue could focus on using machine learning methods to isolate subdomains $D_i \in D$ and construct local estimates of the Lipschitz constant  $L_i$ in the selected subdomains.
-}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Conceptualization and methodology, K.B.; software and validation, I.L., E.K.; formal analysis, K.B.; investigation, I.L.; data curation, I.L.; writing --- original draft preparation, K.B.; writing- -- review and editing, K.B.; visualization, I.L.; funding acquisition, K.B.
+\authorcontributions{Conceptualization and methodology, K.B.; software and validation, A.S., A.Sh.; formal analysis, K.B.; investigation, A.S.; data curation, A.Sh.; writing --- original draft preparation, K.B.; writing- -- review and editing, K.B.; visualization, A.S.; funding acquisition, K.B.
 All authors have read and agreed to the published version of the manuscript.}
@@
 
-%\dataavailability{In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study.
-Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.
-You might choose to exclude this statement if the study did not report any data.}
+\dataavailability{The data analyzed in this study are openly available at https:// .}
 
@@
 \conflictsofinterest{The authors declare no conflict of interest.}
-
-%% Optional
-%\sampleavailability{Samples of the compounds ... are available from the authors.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Only for journal Encyclopedia
-%\entrylink{The Link to this entry published on the encyclopedia platform.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional
-%\abbreviations{Abbreviations}{
-%The following abbreviations are used in this manuscript:\\
-%
-%\noindent 
-%\begin{tabular}{@{}ll}
-%MDPI & Multidisciplinary Digital Publishing Institute\\
-%DOAJ & Directory of open access journals\\
-%TLA & Three letter acronym\\
-%LD & Linear dichroism
-%\end{tabular}}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% Optional
-%\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A").
-If the appendix sections contain a heading then change the argument to "yes".
-%\appendixstart
-%\appendix
-%\section{}
-%\subsection{}
-%The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data.
-Mathematical proofs of results not central to the paper can be added as an appendix.
-%
-%\begin{specialtable}[H] 
-%%\tablesize{\scriptsize}
-%\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab2}}
-%%\tablesize{} % You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
-%\begin{tabular}{ccc}
-%\toprule
-%\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-%\midrule
-%Entry 1		& Data			& Data\\
-%Entry 2		& Data			& Data\\
-%\bottomrule
-%\end{tabular}
-%\end{specialtable}
-%
-%\section{}
-%All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.
 
@@
 
-%=====================================
-% References, variant B: internal bibliography
-%=====================================
-%\begin{thebibliography}{999}
-%% Reference 1
-%\bibitem[Author1(year)]{ref-journal}
-%Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
-%% Reference 2
-%\bibitem[Author2(year)]{ref-book1}
-%Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
-%% Reference 3
-%\bibitem[Author3(year)]{ref-book2}
-%Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
-%% Reference 4
-%\bibitem[Author4(year)]{ref-unpublish}
-%Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} stage of publication (under review; accepted; in~press).
-%% Reference 5
-%\bibitem[Author5(year)]{ref-communication}
-%Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
-%% Reference 6
-%\bibitem[Author6(year)]{ref-proceeding}
-%Author 1, A.B.; Author 2, C.D.; Author 3, E.F.
-Title of Presentation.
-In Title of the Collected Work (if available), Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference; Editor 1, Editor 2, Eds.
-(if available); Publisher: City, Country, Year (if available); Abstract Number (optional), Pagination (optional).
-%% Reference 7
-%\bibitem[Author7(year)]{ref-thesis}
-%Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
-%% Reference 8
-%\bibitem[Author8(year)]{ref-url}
-%Title of Site. Available online: URL (accessed on Day Month Year).
-%\end{thebibliography}
-
-% If authors have biography, please use the format below
-%\section*{Short Biography of Authors}
-%\bio
-%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
-%{\textbf{Firstname Lastname} Biography of first author}
-%
-%\bio
-%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
-%{\textbf{Firstname Lastname} Biography of second author}
-
-% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
-For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2022-02-02 2022/LION/main.tex
LION заготовка

--- 2018-12-28 2019/MOTOR/motor.tex
@@
 %
-\title{ Generalized nested optimization scheme for global optimization problems
-\thanks{This study was supported by the Russian Science Foundation, project 
-No.\,16-11-10150.}}
+\title{ML-based approach for accelerating global search algorithm for solving multicriteria problems\thanks{Supported by organization x.}}
 %
@@
 %
-\author{Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
-Ilia Lebedev\orcidID{0000-0002-8736-0652}}
+\author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
+Vladimir Grishagin\orcidID{0000-0002-2884-3670} \and
+Evgeny Kozinov\orcidID{0000-0001-6776-0096}}
 %
-\authorrunning{K. Barkalov, I. Lebedev}
+\authorrunning{F. Author et al.}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
-\email{konstantin.barkalov@itmm.unn.ru}}
+\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, \email{vagris@unn.ru}}
 %
@@
 The abstract should briefly summarize the contents of the paper in
-150--250 words.
+15--250 words.
 
-\keywords{Dimensionality reduction \and Global optimization \and Multiextremal functions.}
+\keywords{First keyword  \and Second keyword \and Another keyword.}
 \end{abstract}
@@
 \subsection{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is not indented. The first paragraph that follows a table, figure,
+Please note that the first paragraph of a section or subsection is
+not indented. The first paragraph that follows a table, figure,
 equation etc. does not need an indent, either.
 
-Subsequent paragraphs, however, are indented.
+1 \cite{Miettinen1999}
+2 \cite{Ehrgott2005}
+3 \cite{Pardalos2017}
+4 \cite{Strongin2000}
+5 \cite{Sergeyev2013}
+6 \cite{GergelKozinovAPI2016}
+7 \cite{Gergel2018}
+8 \cite{Marler2004}
+9 \cite{GergelKozinov2020}
+10 \cite{Yu2011}
+11 \cite{globalizerSystem}
+12 \cite{Evtushenko2014}
+13 \cite{Zilinskas2015}
+14 \cite{PISA2003}
 
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
-
-%\begin{figure}
-%\includegraphics[width=\textwidth]{fig1.eps}
-%\caption{A figure caption is always placed below the illustration.
-%Please note that short captions are centered, while long ones are
-%justified by the macro package automatically.} \label{fig1}
-%\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{globalizerSystem}, a
-book~\cite{strSergGO}, proceedings without editors~\cite{Jones2009}.
 %
@@
 %
+% \bibliographystyle{splncs04}
+% \bibliography{mybibliography}
+%
 \bibliographystyle{splncs04}
 \bibliography{bibliography}
-%
 
-%\begin{thebibliography}{8}
-%\bibitem{ref_article1}
-%Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-%
-%\bibitem{ref_lncs1}
-%Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-%F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-%Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-%
-%\bibitem{ref_book1}
-%Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-%Location (1999)
-%
-%\bibitem{ref_proc1}
-%Author, A.-B.: Contribution title. In: 9th International Proceedings
-%on Proceedings, pp. 1--2. Publisher, Location (2010)
-%
-%\bibitem{ref_url1}
-%LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-%Oct 2017
-%\end{thebibliography}
 \end{document}

### 2022-02-03 2022/PAVT/PAVT2022.tex
fix doi

@@
 
-As experimental data, information is taken from the article \cite{cao2019}, which represent a change in the concentrations of reaction components over time at different temperatures; as an example of experimental data at a temperature of 276.2 K, the data in the table \ref{table1} are presented.
+As experimental data, information is taken from the article \cite{cao2019}, which represent a change in the concentrations of reaction components over time at different temperatures; as an example of experimental data at a temperature of 276.2 K, the data in Table \ref{table1} are presented.
 
 \begin{table}
+\caption{Experimental Data}
 \label{table1}
-\caption{Experimental Data}
 \begin{center}
@@
 \[
-0=x_1<x_2<...<x_{x+p}=1.
+0=x_1<x_2<...<x_{k+p}=1.
 \]
@@
 Therefore, during the calculations, the value of the upper bound of the pre-exponential multipliers was reduced and the solution was obtained at the upper bounds of $10^5$.
-The kinetic parameters found are presented in the table~\ref{table_res1}.
+The kinetic parameters found are presented in Table~\ref{table_res1}.
 
 \begin{table}
+\caption{Calculated rate constants of reaction}
 \label{table_res1}
-\caption{Calculated rate constants of reaction}
 \begin{center}

### 2022-02-05 2022/LION/main.tex
LION en path1

@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-15--250 words.
+The paper considers a new approach to solving multicriterial optimization problems on the base of criteria scalarization, dimensionality reduction via Peano curves and efficient global search algorithm.
+The novelty of the approach consists in application of machine learning methods combined with utilizing a posteriori information for acceleration of the search.
+Effectiveness of the proposed approach has been demonstrated by means of solving a set of multiextremal multicriterial optimization problems.
 
-\keywords{First keyword  \and Second keyword \and Another keyword.}
+\keywords{Multicriterial Problems \and Global Optimization \and Machine Learning \and Logistic Regression.}
 \end{abstract}
@@
 %
-\section{First Section}
-\subsection{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is
-not indented. The first paragraph that follows a table, figure,
-equation etc. does not need an indent, either.
+\section{Introduction}
+Methods of machine learning (ML) are powerful tools that are applied in many fields of applications.
+In particular, the ML methods are successfully used for solving complicated problems of computational mathematics.
+One of such problem classes to which ML can be applied is the class of multicriterial optimization (MCO) models.
+In this case, ML is used, as a rule, in combination with metaheuristic algorithms \textbf{[links], [Sergeev, Kvasov]} that in the case of multiextremal criteria lose in efficiency to deterministic methods.
+
+Among the most qualitative deterministic methods for solving problems of considered class the information-statistical global search algorithm \cite{Strongin2000,Sergeyev2013} can be considered.
+This method proposed initially for solving scalar problems was successfully extended to MCO models \cite{GergelKozinovAPI2016,Gergel2018}.
+
+The paper reflects results of a new research direction connected with utilizing ML methods for acceleration of the global search algorithm in the case of its application to MCO problems.
+The proposed approach is based on building a number of separation planes in the criteria space to segregate the Pareto set and accelerate the process of its construction.
+The efficiency of the approach is estimated experimentally compared to several other MCO methods.
+
+\section{Problem statement}
+
+The MCO problem to be considered is formulated as follows:
+\begin{equation}
+f(y)=(f_1 (y),f_2 (y),\dots ,f_s (y)) \to \min, y\in D,
+\label{eq:1}
+\end{equation}
+\begin{equation}
+D=\{ y \in R^N: a_i \leq y_i \leq b_i,1 \leq i \leq N,a,b \in R \}.
+\label{eq:2}
+\end{equation}
+where $f_i (y)$, $1 \leq i \leq s$, are  criteria, $y=(y_1,y_2, \dots,y_N)$ is a vector of independent variables, $N$ is the dimension of the problem.
+The criteria $f_i (y)$, $1 \leq i \leq s$, are supposed to be multiextremal, to be given as a ``black-box'' functions and to satisfy the Lipschitz condition
+\begin{equation}
+|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|,y',y'' \in D, 1 \leq i \leq s,
+\label{eq:3}
+\end{equation}
+where $L_i>0$, $1 \leq i \leq s$, are a priori unknown Lipschitz constants.
+
+Any effective (Pareto-optimal) variant  in which it is impossible to reduce the values of all criteria $f_i (y)$, $1 \leq i \leq s$, at once by choosing another option   can be considered as a partial solution to the MCO problem.
+In practice, various scalarization techniques are often used to find effective solutions \cite{Ehrgott2005,Pardalos2017,Marler2004,GergelKozinov2020}.
+The present study uses the minimax convolution of partial criteria that possesses good theoretical properties
+\begin{equation}
+\min{f(y)} \to \min_{y\in D}{\max_{1 \leq i \leq s}{(f_i (y) \lambda_i )}} ,
+\label{eq:4}
+\end{equation}
+where $\lambda_i \geq 0$, $1 \leq i \leq s$, are numerical indicators of the significance of each criterion.
+
+In general case, the entire set of Pareto-optimal variants is taken as a complete solution to the MCO problem.
+For the numerical building of an approximation of the Pareto set a number of scalar optimization problems (\ref{eq:4}) are solved with different coefficients $\lambda_i \geq 0$, $1 \leq i \leq s$.
 

### 2022-02-05 2022/LION/main.tex
add references

@@
 \usepackage{graphicx}
+\usepackage{color}
+\usepackage{hyperref}
+
 % Used for displaying a sample figure. If possible, figure files should
@@
 %
-\authorrunning{F. Author et al.}
+\authorrunning{K. Barkalov, V. Grishagin and E. Kozinov}
 % First names are abbreviated in the running head.
@@
 \section{Introduction}
-Methods of machine learning (ML) are powerful tools that are applied in many fields of applications.
+\textcolor[rgb]{1,0,0}{Machine learning (ML) methods are powerful tools that are applied in many areas of research.}
 In particular, the ML methods are successfully used for solving complicated problems of computational mathematics.
 One of such problem classes to which ML can be applied is the class of multicriterial optimization (MCO) models.
-In this case, ML is used, as a rule, in combination with metaheuristic algorithms \textbf{[links], [Sergeev, Kvasov]} that in the case of multiextremal criteria lose in efficiency to deterministic methods.
+In this case, ML is used, as a rule, in combination with metaheuristic algorithms \cite{Subraveti2019} that in the case of multiextremal criteria lose in efficiency to deterministic methods \cite{Kvasov2018,Sergeyev2018}.
 
-Among the most qualitative deterministic methods for solving problems of considered class the information-statistical global search algorithm \cite{Strongin2000,Sergeyev2013} can be considered.
+Among the most qualitative deterministic methods for solving \textcolor[rgb]{1,0,0}{these problems} the information-statistical global search algorithm \cite{Sergeyev2013,Strongin2000} can be considered.
 This method proposed initially for solving scalar problems was successfully extended to MCO models \cite{GergelKozinovAPI2016,Gergel2018}.
@@
 \begin{equation}
-f(y)=(f_1 (y),f_2 (y),\dots ,f_s (y)) \to \min, y\in D,
+f(y)=(f_1 (y),f_2 (y),\dots ,f_s (y)) \to \min, \; y\in D,
 \label{eq:1}
@@
 \begin{equation}
-D=\{ y \in R^N: a_i \leq y_i \leq b_i,1 \leq i \leq N,a,b \in R \}.
+D=\left\{ y \in R^N: a_i \leq y_i \leq b_i, \; 1 \leq i \leq N, \; a,b \in R \right\}.
 \label{eq:2}
@@
 \begin{equation}
-|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|,y',y'' \in D, 1 \leq i \leq s,
+|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, \; y',y'' \in D, \; 1 \leq i \leq s,
 \label{eq:3}
@@
 
-1 \cite{Miettinen1999}
-2 \cite{Ehrgott2005}
-3 \cite{Pardalos2017}
-4 \cite{Strongin2000}
-5 \cite{Sergeyev2013}
-6 \cite{GergelKozinovAPI2016}
-7 \cite{Gergel2018}
-8 \cite{Marler2004}
-9 \cite{GergelKozinov2020}
-10 \cite{Yu2011}
-11 \cite{globalizerSystem}
-12 \cite{Evtushenko2014}
-13 \cite{Zilinskas2015}
-14 \cite{PISA2003}
+%1 \cite{Miettinen1999}
+%2 \cite{Ehrgott2005}
+%3 \cite{Pardalos2017}
+%4 \cite{Strongin2000}
+%5 \cite{Sergeyev2013}
+%6 \cite{GergelKozinovAPI2016}
+%7 \cite{Gergel2018}
+%8 \cite{Marler2004}
+%9 \cite{GergelKozinov2020}
+%10 \cite{Yu2011}
+%11 \cite{globalizerSystem}
+%12 \cite{Evtushenko2014}
+%13 \cite{Zilinskas2015}
+%14 \cite{PISA2003}
 

### 2022-02-07 2022/LION/main.tex
lambda

@@
 \end{equation}
-where $\lambda_i \geq 0$, $1 \leq i \leq s$, are numerical indicators of the significance of each criterion.
+where $\lambda_i \geq 0$, $1 \leq i \leq s$, $\sum_{i=1}^s{\lambda_i}=1$, are numerical indicators of the significance of each criterion.
 
 In general case, the entire set of Pareto-optimal variants is taken as a complete solution to the MCO problem.
-For the numerical building of an approximation of the Pareto set a number of scalar optimization problems (\ref{eq:4}) are solved with different coefficients $\lambda_i \geq 0$, $1 \leq i \leq s$.
+For the numerical building of an approximation of the Pareto set a number of scalar optimization problems (\ref{eq:4}) are solved with different coefficients $\lambda_i \geq 0$, $1 \leq i \leq s$, uniformly distributed
+%in the space of ones.
+\textcolor[rgb]{1,0,0}{in $[0,1]$.}
 

### 2022-02-07 2022/LION/main.tex
Lion path2

@@
 \usepackage{hyperref}
+
+\usepackage{amsmath}
 
@@
 
+\section{General computational scheme}
+
+In the framework of the proposed approach the information-statistical theory of global search is used for solving the scalar optimization problems (\ref{eq:4}).
+For problems with several variables ($N>1$) the dimensionality reduction on the base of Peano-type mappings \cite{Sergeyev2013,Strongin2000}.
+While solving a problem (\ref{eq:4}) the algorithm generates a sequence of trials where the term ``trial'' means evaluation of criteria at a point of the feasible domain D and computation of the convolution (\ref{eq:4}) at this point.
+
+Coordinates of a new trial are chosen according to the following rules:
+
+\begin{enumerate}
+  \item Obtain univariate images of all points of preceding trials mapped onto one-dimensional interval [0,1] by means of Peano curves.
+  \item Partition the segment [0,1] into subintervals according to the images of trials performed.
+  \item For each subinterval compute a numerical value called \textit{characteristic} of this subinterval.
+  \item Select the subinterval with the highest characteristic.
+  \item Place the point of new trial within the subinterval with the highest characteristic. 
+  \item Map the new trial point to the point $y \in D$ using Peano evolvent and compute the value of the function (\ref{eq:4}) at this point. 
+\end{enumerate}
+
+The algorithm completes execution when the required accuracy is achieved.
+After stopping the global search algorithm, an estimate of the Pareto set is built on the base of the accumulated information obtained in the course of optimization.
+If the quality of the Pareto set estimate is not sufficient, then new preferences (new coefficients $\lambda$) are set and the search process continues.
+
+\section{Approaches to improving search efficiency}
+
+Two techniques are used to improve the search efficiency of the Pareto area assessment.
+
+The first one is to jointly solve a series of global search problems from (\ref{eq:4}).
+The essence of the technique consists in the accumulation of search information during the optimization process and in reusing afterwards this information while solving a problem (\ref{eq:4}) with new weight coefficients $\lambda$.
+A detailed description of the technique can be found in \cite{GergelKozinovAPI2016,Gergel2018}.
+
+The second improvement is connected with a new method for calculating the characteristics of subintervals formed during univariate optimization on the internal [0,1].
+
+Let $R(i)$ be the characteristic of the $i$-th subinterval. This characteristic is supposed to consist of two parts:
+
+\begin{equation}
+R(i)=R_{ags} (i)+ \alpha R_{PD} (i).
+\label{eq:5}
+\end{equation}
+
+The term $R_{ags}(i)$ allows you to select a subinterval oriented at finding the global minimum of the current optimization task (\ref{eq:4}), while $R_{PD}(i)$ influences the selection of a subinterval to improve the evaluation of the Pareto area.
+The coefficient $\alpha$ from (\ref{eq:5}) allows adjusting the contribution of each of parts.
+
+$R_{ags}(i)$ is calculated in accordance with the expression from the global search algorithm \cite{GergelKozinovAPI2016,Gergel2018,Sergeyev2013,Strongin2000}.
+
+Calculation of $R_{PD}(i)$ is based on machine learning methods.
+When solving the first global search problem (\ref{eq:4}), the value of $R_{PD}(i)$ is assumed to be zero.
+To solve each subsequent scalar problem (\ref{eq:4}), $R_{PD}(i)$ is calculated as follows.
+
+In the course of Pareto set assessment all trial points are partitioned into two classes.
+This bipartition is based on belonging to the Pareto set.
+For these classes, a separation hyperplane is constructed in the domain of criteria values using the logistic regression \cite{Yu2011}.
+Examples of separated planes are shown in Fig.
+\ref{fig:1}.
+During the choice of a new trial point, the distances of all obtained criteria values from the separated plane are calculated and the value of $R_{PD}(i)$ is built depending on these distances.
+
+\begin{figure}[ht]
+\includegraphics[width=0.8\textwidth]{fig1.png}
+\caption{Separated hyperplanes in the criteria space} \label{fig:1}
+\end{figure}
+
+Hereinafter the algorithm taking into account distances to separated hyperplane and accumulated information will be called ML\_MGSA.
+
+\section{Results of computational experiments}
+
+Computational testing was carried out on the supercomputer ``Lobachevsky'' of Nizhny Novgorod State University in the environment of the Globalizer software system \cite{globalizerSystem}.
+
+The first series of experiments was performed to compare the algorithm ML\_MGSA with a number of widely known multiobjective optimization algorithms using a test two-criterion problem \cite{Evtushenko2014}
+
+\begin{equation}
+f_1 (y) = (y_1-1) y_2^2+1, f_2 (y) = y_2, 0\leq y_1, y_2 \leq 1.
+\label{eq:6}
+\end{equation}
+
+In these experiments, a numerical approximation of the Pareto set was built.
+The quality of approximation is estimated by such indicators as the hypervolume index (HV) and distribution uniformity index (DU) \cite{Evtushenko2014,Zilinskas2015}.
+In the framework of the experiment 5 algorithms of multicriterial optimization were compared: the Monte-Carlo (MC) method, the genetic algorithm SEMO \cite{PISA2003}, the Non-uniform coverage (NUC) method \cite{Evtushenko2014}, the bi-objective Lipschitz optimization (BLO) method \cite{Zilinskas2015} and algorithm ML\_MGSA, proposed in this article.
+
+ML\_MGSA solved 25 problems (\ref{eq:4}) with different convolution coefficients $\lambda$, uniformly distributed in the field of their variation.
+The results of the conducted experiments are shown in Table \ref{tab:1} where ML\_MGSA with $\alpha=0$ corresponds to the version without machine learning.
+
+\begin{table}[]
+\caption{Efficiency of the multicriterial algorithms compared}
+\label{tab:1}
+\begin{tabular}{ccccccc}
+\hline
+Algorithm                      & MC    & SEMO  & NUC   & BLO   & \begin{tabular}[c]{@{}c@{}}ML\_MGSA \\ ($\alpha= 0$)\end{tabular} & \begin{tabular}[c]{@{}c@{}}ML\_MGSA \\ ($\alpha= 1.5$)\end{tabular} \\ \hline
+Number of trials               & 500   & 500   & 515   & 498   & 302             & 269               \\
+Number of Pareto points found  & 67    & 104   & 29    & 68    & 92              & 79                \\
+HV index (the more the better) & 0.300 & 0.312 & 0.306 & 0.308 & 0.312           & 0.312             \\
+DU index (the less the better) & 1.277 & 1.116 & 0.210 & 0.175 & 0.101           & 0.103             \\ \hline
+\end{tabular}
+\end{table}
+
+As it follows from the results presented in Table \ref{tab:1}, ML\_MGSA demonstrates the better quality compared to other methods.
+
+In the second group of experiments 100 bi-criteria MCO problems have been solved. As the criteria the multiextremal functions
+
+\begin{equation}
+\begin{matrix}
+f(y) = - (AB + CD)^{1/2} \\
+AB = \Big(\sum_{i=1}^7 \sum_{j=1}^7 [A_{ij} a_{ij}(y_1, y_2) + B_{ij} b_{ij}(y_1, y_2)]\Big)^2 \\
+CD = \Big(\sum_{i=1}^7 \sum_{j=1}^7 [C_{ij} a_{ij}(y_1, y_2) - D_{ij} b_{ij}(y_1, y_2)]\Big)^2 \\ 
+a_{ij}(y_1, y_2) = \sin(\pi i y_1) \sin(\pi j y_2), b_{ij}(y_1, y_2) = \cos(\pi i y_1) \cos(\pi j y_2),
+\end{matrix}
+\end{equation}
+$0 \leq y_1,y_2 \leq 1$, have been taken where parameters $-1 \leq A_{ij},B_{ij},C_{ij},D_{ij} \leq 1$ are the independent random numbers distributed uniformly \cite{Grishagin2015_2,Gergel2019_2}.
+
+In order to estimate efficiency of machine learning procedure in ML\_MGSA, this method tested in 2 variants: with ($\alpha>0$) and without ($\alpha=0$) machine learning.
+Each multicriterial problem was reduced to 50 scalar problems (\ref{eq:4}) with different weight coefficients $\lambda$ distributed uniformly.
+Other algorithm's parameters were the same.
+The averaged results can be found in Table \ref{tab:2}.
+
+\begin{table}[]
+\caption{Efficiency of the multicriterial algorithms compared}
+\label{tab:2}
+\begin{tabular}{ccc}
+\hline
+                                  & ML\_MGSA ($\alpha=0.0$) & ML\_MGSA ($\alpha=0.01$) \\ \hline
+Average number of iterations      & 1902.4           & 658.6             \\
+Average value of  DU index        & 1.32             & 1.55              \\
+Average value of HV index         & 92.1             & 91.9              \\
+Reducing the number of iterations & 1                & 2.9     \\  \hline        
+\end{tabular}
+\end{table}
+
+As shown in the table, embedding the machine learning into the global search algorithm results in a significant reduction of the number of trials (almost 3 times) with maintaining close values of indicators HV and DU.
+
 %1 \cite{Miettinen1999}

### 2022-02-07 2022/LION/main.tex
LION PD->PS

@@
 \begin{equation}
-R(i)=R_{ags} (i)+ \alpha R_{PD} (i).
+R(i)=R_{ags} (i)+ \alpha R_{PS} (i).
 \label{eq:5}
@@
 
-The term $R_{ags}(i)$ allows you to select a subinterval oriented at finding the global minimum of the current optimization task (\ref{eq:4}), while $R_{PD}(i)$ influences the selection of a subinterval to improve the evaluation of the Pareto area.
+The term $R_{ags}(i)$ allows you to select a subinterval oriented at finding the global minimum of the current optimization task (\ref{eq:4}), while $R_{PS}(i)$ influences the selection of a subinterval to improve the evaluation of the Pareto area.
 The coefficient $\alpha$ from (\ref{eq:5}) allows adjusting the contribution of each of parts.
@@
 
-Calculation of $R_{PD}(i)$ is based on machine learning methods.
-When solving the first global search problem (\ref{eq:4}), the value of $R_{PD}(i)$ is assumed to be zero.
-To solve each subsequent scalar problem (\ref{eq:4}), $R_{PD}(i)$ is calculated as follows.
+Calculation of $R_{PS}(i)$ is based on machine learning methods.
+When solving the first global search problem (\ref{eq:4}), the value of $R_{PS}(i)$ is assumed to be zero.
+To solve each subsequent scalar problem (\ref{eq:4}), $R_{PS}(i)$ is calculated as follows.
 
@@
 \ref{fig:1}.
-During the choice of a new trial point, the distances of all obtained criteria values from the separated plane are calculated and the value of $R_{PD}(i)$ is built depending on these distances.
+During the choice of a new trial point, the distances of all obtained criteria values from the separated plane are calculated and the value of $R_{PS}(i)$ is built depending on these distances.
 

### 2022-02-07 2022/LION/main.tex
fix some typos

@@
 For problems with several variables ($N>1$) the dimensionality reduction on the base of Peano-type mappings \cite{Sergeyev2013,Strongin2000}.
-While solving a problem (\ref{eq:4}) the algorithm generates a sequence of trials where the term ``trial'' means evaluation of criteria at a point of the feasible domain D and computation of the convolution (\ref{eq:4}) at this point.
+While solving a problem (\ref{eq:4}) the algorithm generates a sequence of trials where the term ``trial'' means evaluation of criteria at a point of the feasible domain $D$ and computation of the convolution (\ref{eq:4}) at this point.
 
@@
 \begin{enumerate}
-  \item Obtain univariate images of all points of preceding trials mapped onto one-dimensional interval [0,1] by means of Peano curves.
-  \item Partition the segment [0,1] into subintervals according to the images of trials performed.
+  \item Obtain univariate images of all points of preceding trials mapped onto one-dimensional interval $[0,1]$ by means of Peano curves.
+  \item Partition the segment $[0,1]$ into subintervals according to the images of trials performed.
   \item For each subinterval compute a numerical value called \textit{characteristic} of this subinterval.
@@
 
-The second improvement is connected with a new method for calculating the characteristics of subintervals formed during univariate optimization on the internal [0,1].
+The second improvement is connected with a new method for calculating the characteristics of subintervals formed during univariate optimization on the internal $[0,1]$.
 
 Let $R(i)$ be the characteristic of the $i$-th subinterval. This characteristic is supposed to consist of two parts:
-
 \begin{equation}
@@
 For these classes, a separation hyperplane is constructed in the domain of criteria values using the logistic regression \cite{Yu2011}.
-Examples of separated planes are shown in Fig.
-\ref{fig:1}.
+Examples of separated planes are shown in Fig.~\ref{fig:1}.
 During the choice of a new trial point, the distances of all obtained criteria values from the separated plane are calculated and the value of $R_{PS}(i)$ is built depending on these distances.
@@
 \begin{equation}
-f_1 (y) = (y_1-1) y_2^2+1, f_2 (y) = y_2, 0\leq y_1, y_2 \leq 1.
+f_1 (y) = (y_1-1) y_2^2+1, \; f_2 (y) = y_2, \; 0\leq y_1, y_2 \leq 1.
 \label{eq:6}
@@
 
-\begin{table}[]
+\begin{table}
 \caption{Efficiency of the multicriterial algorithms compared}
@@
 \end{equation}
-$0 \leq y_1,y_2 \leq 1$, have been taken where parameters $-1 \leq A_{ij},B_{ij},C_{ij},D_{ij} \leq 1$ are the independent random numbers distributed uniformly \cite{Grishagin2015_2,Gergel2019_2}.
+$0 \leq y_1,y_2 \leq 1$, have been taken where parameters $-1 \leq A_{ij},B_{ij},C_{ij},D_{ij} \leq 1$ are the independent random numbers distributed uniformly \cite{Gergel2019_2,Grishagin2015_2}.
 

### 2022-02-08 2022/LION/main.tex
fix figures in LaTeX

@@
 \usepackage{hyperref}
+\usepackage{ulem}
 
@@
 %
-\title{ML-based approach for accelerating global search algorithm for solving multicriteria problems\thanks{Supported by organization x.}}
+\title{ML-based approach for accelerating global search algorithm for solving multicriteria problems\thanks{This study was supported by the Russian Science Foundation, project No 21-11-00204.}}
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{ML-based approach for accelerating global search algorithm}
 % If the paper title is too long for the running head, you can set
@@
 \section{Introduction}
-\textcolor[rgb]{1,0,0}{Machine learning (ML) methods are powerful tools that are applied in many areas of research.}
+Machine learning (ML) methods are powerful tools that are applied in many areas of research.
 In particular, the ML methods are successfully used for solving complicated problems of computational mathematics.
@@
 
-Among the most qualitative deterministic methods for solving \textcolor[rgb]{1,0,0}{these problems} the information-statistical global search algorithm \cite{Sergeyev2013,Strongin2000} can be considered.
+Among the most qualitative deterministic methods for solving these problems the information-statistical global search algorithm \cite{Sergeyev2013,Strongin2000} can be considered.
 This method proposed initially for solving scalar problems was successfully extended to MCO models \cite{GergelKozinovAPI2016,Gergel2018}.
@@
 The MCO problem to be considered is formulated as follows:
-\begin{equation}
-f(y)=(f_1 (y),f_2 (y),\dots ,f_s (y)) \to \min, \; y\in D,
-\label{eq:1}
-\end{equation}
-\begin{equation}
-D=\left\{ y \in R^N: a_i \leq y_i \leq b_i, \; 1 \leq i \leq N, \; a,b \in R \right\}.
-\label{eq:2}
-\end{equation}
+\begin{gather}
+f(y)=(f_1 (y),f_2 (y),\dots ,f_s (y)) \to \min, \; y\in D,  \label{eq:1} \\ 
+D=\left\{ y \in R^N: a_i \leq y_i \leq b_i, \; 1 \leq i \leq N, \; a,b \in R \right\}.\label{eq:2} 
+\end{gather}
 where $f_i (y)$, $1 \leq i \leq s$, are  criteria, $y=(y_1,y_2, \dots,y_N)$ is a vector of independent variables, $N$ is the dimension of the problem.
@@
 Any effective (Pareto-optimal) variant  in which it is impossible to reduce the values of all criteria $f_i (y)$, $1 \leq i \leq s$, at once by choosing another option   can be considered as a partial solution to the MCO problem.
-In practice, various scalarization techniques are often used to find effective solutions \cite{Ehrgott2005,Pardalos2017,Marler2004,GergelKozinov2020}.
+In practice, various scalarization techniques are often used to find effective solutions \cite{Ehrgott2005,GergelKozinov2020,Marler2004,Pardalos2017}.
 The present study uses the minimax convolution of partial criteria that possesses good theoretical properties
@@
 In general case, the entire set of Pareto-optimal variants is taken as a complete solution to the MCO problem.
-For the numerical building of an approximation of the Pareto set a number of scalar optimization problems (\ref{eq:4}) are solved with different coefficients $\lambda_i \geq 0$, $1 \leq i \leq s$, uniformly distributed
+For the numerical building of an approximation of the Pareto set a number of scalar optimization problems (\ref{eq:4}) are solved with different coefficients $\lambda_i \geq 0$, $1 \leq i \leq s$, distributed uniformly.
 %in the space of ones.
-\textcolor[rgb]{1,0,0}{in $[0,1]$.}
+%\textcolor[rgb]{1,0,0}{in $[0,1]$.}
 
@@
 For these classes, a separation hyperplane is constructed in the domain of criteria values using the logistic regression \cite{Yu2011}.
-Examples of separated planes are shown in Fig.~\ref{fig:1}.
-During the choice of a new trial point, the distances of all obtained criteria values from the separated plane are calculated and the value of $R_{PS}(i)$ is built depending on these distances.
+Examples of separating planes are shown in Fig.~\ref{fig:1}.
+During the choice of a new trial point, the distances of all obtained criteria values from the separating plane are calculated and the value of $R_{PS}(i)$ is built depending on these distances.
 
 \begin{figure}[ht]
-\includegraphics[width=0.8\textwidth]{fig1.png}
-\caption{Separated hyperplanes in the criteria space} \label{fig:1}
+\center
+%\includegraphics[width=0.9\textwidth]{fig1.png}
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=0.9\linewidth]{fig1a.png}}
+\end{minipage}
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=0.9\linewidth]{fig1b.png}}
+\end{minipage}
+
+\caption{Separating hyperplanes in the criteria space} \label{fig:1}
 \end{figure}
 
-Hereinafter the algorithm taking into account distances to separated hyperplane and accumulated information will be called ML\_MGSA.
+Hereinafter the algorithm taking into account distances to separating hyperplane and accumulated information will be called ML\_MGSA.
 
@@
 
-Computational testing was carried out on the supercomputer ``Lobachevsky'' of Nizhny Novgorod State University in the environment of the Globalizer software system \cite{globalizerSystem}.
+Computational testing was carried out on the``Lobachevsky'' supercomputer of Nizhny Novgorod State University in the environment of the Globalizer software system \cite{globalizerSystem}.
 
@@
 
-ML\_MGSA solved 25 problems (\ref{eq:4}) with different convolution coefficients $\lambda$, uniformly distributed in the field of their variation.
+ML\_MGSA solved 25 problems (\ref{eq:4}) with different convolution coefficients $\lambda$, distributed uniformly.
+%in the field of their variation.
 The results of the conducted experiments are shown in Table \ref{tab:1} where ML\_MGSA with $\alpha=0$ corresponds to the version without machine learning.
+As it follows from the results presented in Table \ref{tab:1}, ML\_MGSA demonstrates the better quality compared to other methods.
 
@@
 
-As it follows from the results presented in Table \ref{tab:1}, ML\_MGSA demonstrates the better quality compared to other methods.
-
 In the second group of experiments 100 bi-criteria MCO problems have been solved. As the criteria the multiextremal functions
@@
 
-In order to estimate efficiency of machine learning procedure in ML\_MGSA, this method tested in 2 variants: with ($\alpha>0$) and without ($\alpha=0$) machine learning.
-Each multicriterial problem was reduced to 50 scalar problems (\ref{eq:4}) with different weight coefficients $\lambda$ distributed uniformly.
+In order to estimate efficiency of machine learning procedure in ML\_MGSA, this method tested in 2 variants: with ($\alpha>0$) and without ($\alpha=0$)
+\sout{machine learning} \textcolor[rgb]{1,0,0}{use of separating planes in the criteria space}. 
+Each multicriterial problem was reduced to 50 scalar problems (\ref{eq:4}) with different weight coefficients $\lambda$, distributed uniformly.
 Other algorithm's parameters were the same.
@@
 
-%1 \cite{Miettinen1999}
-%2 \cite{Ehrgott2005}
-%3 \cite{Pardalos2017}
-%4 \cite{Strongin2000}
-%5 \cite{Sergeyev2013}
-%6 \cite{GergelKozinovAPI2016}
-%7 \cite{Gergel2018}
-%8 \cite{Marler2004}
-%9 \cite{GergelKozinov2020}
-%10 \cite{Yu2011}
-%11 \cite{globalizerSystem}
-%12 \cite{Evtushenko2014}
-%13 \cite{Zilinskas2015}
-%14 \cite{PISA2003}
-
 %

### 2022-02-10 2022/PAVT/PAVT2022.tex
fix reviewer comment in the UNN part

@@
 \usepackage{cite}
+\usepackage{color}
 
@@
 Then, with the found rate constants of the reaction stages, the system of differential equations (\ref{eq:one})--(\ref{eq:twelve}) is solved, then the calculated concentrations are compared with the corresponding experimental data.
-Mathematically, this problem has the following formulation: it is necessary to minimize the following objective function:
+Mathematically, this problem has the following formulation: it is necessary to minimize the following objective function
 
@@
 
-As it has been already mentioned above, from the formal point of view, we consider the inverse problem of chemical kinetics as Lipschitz global optimization problem.
+\textcolor[rgb]{1,0,0}{As it has been already mentioned above, from the formal point of view we consider the inverse problem of chemical kinetics as a global optimization problem. 
+In the specific problem under consideration, the values of the objective function are calculated by solving stiff ODE system (\ref{eq:one})--(\ref{eq:twelve}).
+Since the right parts of the system are continuous functions with bounded derivatives, then theoretically its solution will also be continuous and bounded.
+Therefore, the discrepancy (\ref{eq:objective_func2}) will satisfy the Lipschitz condition with a priori unknown constant.}
+
 In the general form, a problem of the class specified above can be formulated mathematically as 
@@
 So far, a search trial at some point $x'\in[0,1]$ will include first the construction of the image $y'=y(x')$ and then the computing the value of the function $z'=\varphi(y')$.
+
+%Данный комментарий можно и не включать...
+%Note that the presence of rapidly decreasing components in the solution of the stiff ODE system (\ref{eq:one})--(\ref{eq:twelve}) can lead to the fact that in different subdomains the objective function can have different local Lipschitz constants.
+This problem will affect the convergence of any of the global optimization methods.
+To take it into account, a number of approaches have been proposed related to local tuning of global optimization methods (see, e.g., \cite{Barkalov2021,Strongin2020}).
 
@@
 The asynchronous global optimization algorithm was implemented using C++ (GCC 5.5.0 and Intel MPI were used); the objective function values were computed using Python 3.9.
+\textcolor[rgb]{1,0,0}{The accuracy of solving the ODE system (\ref{eq:one})--(\ref{eq:twelve}) was set small enough so that the final error of the solution was much less than the accuracy of the stopping criterion of the optimization method and did not affect the method used.}
 
@@
 
+\textcolor[rgb]{1,0,0}{Note that when solving the global optimization problem, the number of iterations of the parallel algorithm (and hence its speedup) depends significantly on the estimation of the Lipschitz constant of the objective function.
+The constant is adaptively estimated during the work of the algorithm and may vary depending on the accumulated search information.
+With a correct estimate of the Lipschitz constant the method can converge to a global optimum point faster than in the case of its incorrect estimate.
+This explains the effect of superlinear speedup observed in experiments.}
+
 \begin{table}

### 2022-02-11 2022/PAVTOneAPI/OneAPI.tex
Исправлена часть замечаний

--- 2021-12-15 2022/PAVTOneAPI/OneAPI.tex
@@
 
-\title{Решение многомерных задач глобальной оптимизации с использованием Intel OneApi\footnote{На что ссылаемся?}}
+\title{Решение многомерных задач глобальной оптимизации с использованием Intel oneAPI\footnote{Исследование выполнено за счет гранта Российского научного фонда № 21-11-00204.}}
 
@@
 В рамках данного подхода решение многомерной задачи сводится к решению одномерной.
-Для это используется редукция размерности с помощью кривой Пеано \cite{Sergeyev2013}., которая непрерывно и однозначно отображает отрезок вещественной оси \([0,1]\) на  \(N\)-мерный куб.
+Для этого используется редукция размерности с помощью кривой Пеано \(y(x)\), которая непрерывно и однозначно отображает отрезок вещественной оси \([0,1]\) на \(N\)-мерный куб \cite{Sergeyev2013}.
 
-В процессе работы алгоритма строится последовательность точек \(x^k\), где в каждой точке вычисляется значение минимизируемой функции (производится испытание) \(y^k=f(x^k)\)  Параллельность обеспечиваться следующим образом: в ходе выполнения одной итерации алгоритма будет проводиться \(P\) испытаний одновременно, где \(P \geq 1\).
+В процессе работы алгоритма строится последовательность точек \(x^k\), в которых проводятся \textit{испытания}, т.е.
+вычисляются значения целевой функции \(\phi^k=\phi(y(x^k))\).
+Распараллеливание организовано следующим образом: в ходе выполнения одной итерации алгоритма одновременно проводятся \(P \geq 1\) испытаний в различных точках области поиска.
 
-Одним из вариантов реализации данного подхода является использование инструментов Intel OneAPI.
-Он прост, открыт и легко интегрируя в существующий код.
-Intel OneAPI позволяет написать код один раз и в дальнейшем запускать его на различных устройствах.
+Одним из вариантов реализации данного подхода является использование инструментов Intel oneAPI.
+Он прост, открыт и легко интегрируется в существующий код.
+Intel oneAPI позволяет написать код один раз и в дальнейшем запускать его на различных устройствах.
+В случае данной задачи используем возможности языка \textit{data parallel c++}.
 
-На данный момент получены результаты вычислительных экспериментов на генераторе тестовых задач GKLS \cite{GKLS}.
-В таблице \ref{table:GKLS_RES} приведены среднее число итераций параллельного алгоритма глобального поиска и ускорение относительно последовательной версии алгоритма.
-В каждом эксперименте решалось 100 размерности 4 и 5.
+На данный момент получены результаты вычислительных экспериментов, где тестовые задачи формируются GKLS-генератором \cite{GKLS}.
+В таблице \ref{table:GKLS_RES} приведены среднее число итераций параллельного алгоритма глобального поиска и ускорение относительно последовательной версии.
+В каждом эксперименте решалось 100 задач размерности \(N = 4\) и \(N = 5\).
 
@@
     \hline
-    P    & Итераций 4D & ускорение 4D &         Итераций 5D & ускорение 5D \\ \hline
-	128 & 106,46   & 4,73      &         235,64   & 4,37      \\ \hline
-	256 & 41,25    & 8,50      &         90,45    & 9,20      \\ \hline
+    P    & Итерации (N=4) & Ускорение (N=4) &         Итерации (N=5) & Ускорение (N=5) \\ \hline
+	128 & 106.5   & 4.7      &         235.6   & 4.4      \\ \hline
+	256 & 41.3    & 8.5      &         90.5    & 9.2      \\ \hline
 	\end{tabular}
@@
 2013, 280с.
+
+\bibitem{Sergeyev2013}
+Sergeyev, Y.D., Strongin, R.G., Lera, D. Introduction to global optimization exploiting space-filling curves. 
+New York: Springer, 2013. 125 p.
 

### 2022-02-14 2022/LION/main.tex
submitted

@@
 The paper reflects results of a new research direction connected with utilizing ML methods for acceleration of the global search algorithm in the case of its application to MCO problems.
-The proposed approach is based on building a number of separation planes in the criteria space to segregate the Pareto set and accelerate the process of its construction.
+The proposed approach is based on building a number of separating planes in the criteria space to segregate the Pareto set and accelerate the process of its construction.
 The efficiency of the approach is estimated experimentally compared to several other MCO methods.
@@
 This bipartition is based on belonging to the Pareto set.
-For these classes, a separation hyperplane is constructed in the domain of criteria values using the logistic regression \cite{Yu2011}.
+For these classes, a separating hyperplane is constructed in the domain of criteria values using the logistic regression \cite{Yu2011}.
 Examples of separating planes are shown in Fig.~\ref{fig:1}.
@@
 In order to estimate efficiency of machine learning procedure in ML\_MGSA, this method tested in 2 variants: with ($\alpha>0$) and without ($\alpha=0$)
-\sout{machine learning} \textcolor[rgb]{1,0,0}{use of separating planes in the criteria space}. 
+use of separating planes in the criteria space. 
 Each multicriterial problem was reduced to 50 scalar problems (\ref{eq:4}) with different weight coefficients $\lambda$, distributed uniformly.

### 2022-02-14 2022/PAVTOneAPI/OneAPI.tex
Исправлено очередное замечание.
Еще одно исправление.
fix some misprints

@@
 
-\authors{К.А. Баркалов\superscript{1}, И.Г. Лебедев\superscript{1} Я.В. Кольтюшкина\superscript{1}}
+\authors{К.А. Баркалов\superscript{1}, И.Г. Лебедев\superscript{1}, Я.В. Кольтюшкина\superscript{1}}
 \organizations{Нижегородский государственный университет им. Н.И. Лобачевского\superscript{1}}
@@
 Задача многомерной многоэкстремальной оптимизации может быть определена как поиск наименьшего значения действительной функции \(\phi(y)\)  в гиперинтервале \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
-Следует отметить, что в общем случае задача задана по принципу черного ящика. Для целевой функции предполагаем выполнение условия Липшица с априори неизвестной константой \(L\).
+Будем предполагать, что целевая функция задана как <<черный ящик>> и удовлетворяет условию Липшица с априори неизвестной константой \(L\).
 
@@
 Стронгина разработан эффективный подход к решению задач глобальной оптимизации \cite{Strongin2013}.
-В рамках данного подхода решение многомерной задачи сводится к решению одномерной.
+В рамках данного подхода решение многомерной задачи сводится к решению эквивалентной ей одномерной задачи.
 Для этого используется редукция размерности с помощью кривой Пеано \(y(x)\), которая непрерывно и однозначно отображает отрезок вещественной оси \([0,1]\) на \(N\)-мерный куб \cite{Sergeyev2013}.
@@
 
-Одним из вариантов реализации данного подхода является использование инструментов Intel oneAPI.
-Он прост, открыт и легко интегрируется в существующий код.
+Одним из вариантов реализации данного подхода является использование набора инструментов Intel oneAPI.
+Достоинством Intel oneAPI является простота, открытость и возможность легкой интегрирации в существующий код.
 Intel oneAPI позволяет написать код один раз и в дальнейшем запускать его на различных устройствах.
-В случае данной задачи используем возможности языка \textit{data parallel c++}.
 
-На данный момент получены результаты вычислительных экспериментов, где тестовые задачи формируются GKLS-генератором \cite{GKLS}.
+На данный момент получены результаты вычислительных экспериментов, в которых тестовые задачи формируются специальным генератором \cite{GKLS}.
 В таблице \ref{table:GKLS_RES} приведены среднее число итераций параллельного алгоритма глобального поиска и ускорение относительно последовательной версии.
 В каждом эксперименте решалось 100 задач размерности \(N = 4\) и \(N = 5\).
-
+Эксперименты проведены на компьютере с процессором Intel Core I5-7300HQ 2.5 GHz с видеокартой HD Graphics 630, 16 GB RAM, использовался язык \textit{data parallel c++}.
 \begin{table}[!hbp]
     \centering
-    \caption{Число итераций и ускорение при решении серии 4-х и 5-ти мерных задач}
+    \caption{Число итераций и ускорение при решении серии задач размерности $N=4,5$}
     \begin{tabular}{|c|c|c|c|c|}
     \hline
-    P    & Итерации (N=4) & Ускорение (N=4) &         Итерации (N=5) & Ускорение (N=5) \\ \hline
+    P    & Итерации ($N=4$) & Ускорение ($N=4$) &         Итерации ($N=5$) & Ускорение ($N=5$) \\ \hline
 	128 & 106.5   & 4.7      &         235.6   & 4.4      \\ \hline
@@
 \bibitem{Strongin2013}
-Стронгин Р.Г. Гергель В.П. Гришагин В.А. Баркалов К.А. Параллельные вычисления
-в задачах глобальной оптимизации.
-Москва: Издательство Московского университета,
-2013, 280с.
+Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А. Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство МГУ, 2013. 280 с.
 
 \bibitem{Sergeyev2013}
-Sergeyev, Y.D., Strongin, R.G., Lera, D. Introduction to global optimization exploiting space-filling curves. 
-New York: Springer, 2013. 125 p.
+Sergeyev Y.D., Strongin R.G., Lera D. Introduction to global optimization exploiting space-filling curves. New York: Springer, 2013. 125 p.
 
 \bibitem{GKLS}
-Gaviano M.
+Сергеев Я.Д., Квасов Д.Е. Диагональные методы глобальной оптимизации. М.: ФИЗМАТЛИТ, 2008. 352 с.
+%Gaviano M.
 Lera D.

### 2022-02-15 2022/PAVTOneAPI/OneAPI.tex
fix grant

@@
 
-\title{Решение многомерных задач глобальной оптимизации с использованием Intel oneAPI\footnote{Исследование выполнено за счет гранта Российского научного фонда № 21-11-00204.}}
+\title{Решение многомерных задач глобальной оптимизации с использованием Intel oneAPI\footnote{Работа выполнена при поддержке МинОбрНауки (проект № 0729-2020-0055) и научно-образовательного математического центра <<Математика технологий будущего>> (проект № 075-02-2020-1483/1).}}
 

### 2022-02-28 2022/PAVT/PAVT2022.tex
corrections from Ufa

@@
 
-\textcolor[rgb]{1,0,0}{As it has been already mentioned above, from the formal point of view we consider the inverse problem of chemical kinetics as a global optimization problem. 
+As it has been already mentioned above, from the formal point of view we consider the inverse problem of chemical kinetics as a global optimization problem. 
 In the specific problem under consideration, the values of the objective function are calculated by solving stiff ODE system (\ref{eq:one})--(\ref{eq:twelve}).
 Since the right parts of the system are continuous functions with bounded derivatives, then theoretically its solution will also be continuous and bounded.
-Therefore, the discrepancy (\ref{eq:objective_func2}) will satisfy the Lipschitz condition with a priori unknown constant.}
+Therefore, the discrepancy (\ref{eq:objective_func2}) will satisfy the Lipschitz condition with a priori unknown constant.
 
@@
 The asynchronous global optimization algorithm was implemented using C++ (GCC 5.5.0 and Intel MPI were used); the objective function values were computed using Python 3.9.
-\textcolor[rgb]{1,0,0}{The accuracy of solving the ODE system (\ref{eq:one})--(\ref{eq:twelve}) was set small enough so that the final error of the solution was much less than the accuracy of the stopping criterion of the optimization method and did not affect the method used.}
+The accuracy of solving the ODE system (\ref{eq:one})--(\ref{eq:twelve}) was set small enough so that the final error of the solution was much less than the accuracy of the stopping criterion of the optimization method and did not affect the method used.
 
@@
 
-\textcolor[rgb]{1,0,0}{Note that when solving the global optimization problem, the number of iterations of the parallel algorithm (and hence its speedup) depends significantly on the estimation of the Lipschitz constant of the objective function.
+Note that when solving the global optimization problem, the number of iterations of the parallel algorithm (and hence its speedup) depends significantly on the estimation of the Lipschitz constant of the objective function.
 The constant is adaptively estimated during the work of the algorithm and may vary depending on the accumulated search information.
 With a correct estimate of the Lipschitz constant the method can converge to a global optimum point faster than in the case of its incorrect estimate.
-This explains the effect of superlinear speedup observed in experiments.}
+This explains the effect of superlinear speedup observed in experiments.
 
@@
 So, for example, for the first stage of the reaction $E_1 = 24.65$ kJ/mol, $k_1^0 = 6.85 \cdot 10^4$ min$^{-1}$, for the third stage $E_3 = 75.57$ kJ/mol, $k_3^0 = 2.62 \cdot 10^{14}$ kg $\cdot$mol$^{-1}\cdot$min$^{-1}$.
-However, for some stages, the parameter values were fairly high $k^0$, namely for the second stage $k_2^0 = 8.52 \cdot 10^{25}$ min$^{-1}$, for the seventh stage $k_7^0 = 2.82 \cdot 10^{26}$ kg mol$^{-2}$ min$^{-2}$.
+However, for some stages, the parameter values $k^0$ were fairly high, namely for the second stage $k_2^0 = 8.52 \cdot 10^{25}$ min$^{-1}$, for the seventh stage $k_7^0 = 2.82 \cdot 10^{26}$ kg mol$^{-2}$ min$^{-2}$.
 Therefore, in future works, the constants corresponding to these kinetic parameters will be recalculated.
+Since the problem is multidimensional, there may be several solutions that describe experimental data within a certain deviation error.
+The solution may change if, for example, additional experimental data are found.
+However, the algorithm proposed in the paper provides a single solution within the search range, and high values of pre-exponential multipliers explain the presence of complex stages, which in turn consist of several stages that can be refined by chemists.
 

### 2022-03-02 2022/Sensors/sensors.tex
first virsion of figures
Title and Abstract

@@
 % Full title of the paper (Capitalized)
-\Title{Title}
+\Title{A Fast kNN Algorithm Using Multiple Space-Filling Curves}
 		   %Acceleration of Global Optimization Algorithm by Detecting Local Extrema Based on Machine Learning
@@
 \abstract{
-abstract
+The paper considers a time-efficient implementation of the k-nearest neighborhood (kNN) algorithm.
+A well-known approach for accelerating the kNN algorithm is to utilize dimensionality reduction methods based on the use of space-filling curves.
+In this paper, we take this approach further and propose an algorithm that employs multiple space-filling curves and is faster (with comparable quality) compared to the kNN algorithm which uses kd-trees to determine the nearest neighbors.
+A specific method for constructing multiple Peano curves is outlined, and statements are given about the preservation of object proximity information in the course of dimensionality reduction.
+An experimental comparison with known kNN implementations using kd-trees was performed using test and real-life data.
 }
@@
 \begin{figure}
-\includegraphics[width=0.6\linewidth]{Definitions/logo-mdpi}
+\center
+\includegraphics[width=0.5\linewidth]{fig1.png}
 \caption{Using GSA-DT algorithm to solve GKLS problems}
-\label{fig4}
+\label{fig1}
 \end{figure}
@@
 \end{paracol}
+
 \begin{figure}	
 \widefigure
-\includegraphics[width=15 cm]{Definitions/logo-mdpi}
+\includegraphics[width=1.0\linewidth]{fig2.png}
 \caption{This is a wide figure.\label{fig2}}
 \end{figure}
+
 \begin{paracol}{2}

### 2022-03-16 2022/PAVTOneAPI/PAVT_Poster/poster.tex
Первая версия постера на ПАВТ

\documentclass{beamer}
\usepackage[orientation=portrait, size=a1, scale=1.4, debug]{beamerposter}
\usetheme{unn}

\usepackage{cmbright}

\usepackage{fontspec}

\setmainfont{Times New Roman}
\setromanfont{Times New Roman}
\setsansfont{Times New Roman}

\usepackage{polyglossia}
\setmainlanguage{russian}
\usepackage{csquotes}

\usepackage{booktabs}
\usepackage{ragged2e}

\usepackage[backend=biber,
            movenames=false,
            maxnames=4,
            style=gost-numeric,
            sorting=nty,
            autolang=other]{biblatex}

%\newfontfamily\cyrillicfonttt{lmmonolt10-regular.otf}
%\newfontfamily\cyrillicfontsf{lmsans10-regular.otf}

\addbibresource{bibliography.bib}

\DeclareMathOperator{\re}{\operatorname{Re}}

\title{Решение многомерных задач глобальной оптимизации с использованием Intel oneAPI }
\author{К.А. Баркалов \and И.Г. Лебедев \and Я.В. Кольтюшкина}
\institute{Нижегородский государственный университет им. Н.И. Лобачевского}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}

\begin{document}
\begin{frame}[t]
    \begin{columns}[t]
        \begin{column}[t]{0.48\paperwidth}
            \begin{block}{Постановка задачи}
             Задача многомерной многоэкстремальной оптимизации может быть определена как поиск наименьшего значения действительной функции \(\phi(y)\)  в гиперинтервале \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\). 
Будем предполагать, что целевая функция задана как <<черный ящик>> и удовлетворяет условию Липшица с априори неизвестной константой \(L\)

          \end{block}

          \begin{block}{Редукция размерности с помощью кривых Пеано}

В ННГУ им. Н.И. Лобачевского под руководством проф. Р.Г. Стронгина разработан эффективный подход к решению задач глобальной оптимизации \cite{Strongin2013}. В рамках данного подхода решение многомерной задачи сводится к решению эквивалентной ей одномерной задачи. Для этого используется редукция размерности с помощью кривой Пеано \(y(x)\), которая непрерывно и однозначно отображает отрезок вещественной оси \([0,1]\) на \(N\)-мерный куб \cite{Sergeyev2013}.

 \begin{minipage}[t]{.48\textwidth}
              \begin{figure}
                  \centering
                  \includegraphics[scale=1.27]{images/pean_2.jpg}
              \end{figure}
              \end{minipage}

\end{block}
\begin{block}{Метод глобальной оптимизации}

В процессе работы алгоритма строится последовательность точек \(x^k\), в которых проводятся \textit{испытания}, т.е. вычисляются значения целевой функции \(\phi^k=\phi(y(x^k))\).

Общая схема одной итерации одномерного метода:
              \begin{enumerate}
                \justifying
                \item Упорядочить точки предшествующих испытаний в порядке возрастания их координат: \(a=x_{0}<...<x_{i}<...<x_{k}=b\).
                \item Вычислить для каждого интервала \((x_{i-1};x_{i}),1\leq i\leq k\)  характеристику \(R(i)\) .
                \item Определить интервал \((x_{t-1};x_{t})\) , которому соответствует максимальная характеристика \(R(t)=\max\{R(i),1\leq i\leq k\}\).
                \item Провести следующее испытание в точке \(x^{k+1}=d(t)\in (x_{t-1};x_{t})\) , где \(d(t)\)  — правило размещения точки следующего испытания в интервале с номером \(t\).
                \item Проверить выполнение критерия остановки \(x_{t}-x_{t-1}<\varepsilon\).
              \end{enumerate}

 \begin{minipage}[t]{.48\textwidth}
              \begin{figure}
                  \centering
                  \includegraphics[scale=1.27]{images/PlotGS.png}
              \end{figure}
              \end{minipage}
\end{block}

        \end{column}
        \begin{column}[t]{0.48\paperwidth}
          \begin{block}{Организация распараллеливания}

Распараллеливание организовано следующим образом: в ходе выполнения одной итерации алгоритма одновременно проводятся \(P \geq 1\) испытаний в различных точках области поиска. Важно отметить, что рассматриваются функции, в которых самым трудоемким и времязатратным этапом работы является это вычисление значения функции в точке. Поэтому организация одновременных вычислений этих значений положительно влияет на скорость работы алгоритма для таких функций.

\end{block}

\begin{block}{Использование Intel oneAPI}

Одним из вариантов реализации данного подхода является использование набора инструментов Intel oneAPI. Достоинством Intel oneAPI является простота, открытость и возможность легкой интегрирации в существующий код. Intel oneAPI позволяет написать код один раз и в дальнейшем запускать его на различных устройствах.

В рамках данной задачи нам необходимы возможности распараллеливания, которые обеспечиваются за счет включения в oneAPI языка data parallel c++ и набора библиотек, облегчающих межархитектурную разработку.

\end{block}
%----------------------------------------------------------------------------------------
%	RESULTS
%----------------------------------------------------------------------------------------

\begin{block}{Результаты}

На данный момент получены результаты вычислительных экспериментов, в которых тестовые задачи формируются специальным генератором \cite{GKLS}. В таблицах \ref{table:GKLS_RES_1}, \ref{table:GKLS_RES_2} приведены среднее число итераций параллельного алгоритма глобального поиска и ускорение относительно последовательной версии. В каждом эксперименте решалось 100 задач размерности \(N = 4\) и \(N = 5\). Эксперименты проведены на компьютере с процессором Intel Core I5-7300HQ 2.5 GHz с видеокартой HD Graphics 630, 16 GB RAM, использовался язык \textit{data parallel c++}.

\begin{table}[!hbp]
    \centering
    \caption{Число итераций и ускорение при решении серии задач размерности $N=4$}
    \begin{tabular}{|c|c|c|}
    \hline
    P    & Итерации ($N=4$) & Ускорение ($N=4$) \\ \hline
	128 & 106.5   & 4.7    \\ \hline
	256 & 41.3    & 8.5          \\ \hline
	\end{tabular}

    \label{table:GKLS_RES_1}
\end{table}

\begin{table}[!hbp]
    \centering
    \caption{Число итераций и ускорение при решении серии задач размерности $N=5$}
    \begin{tabular}{|c|c|c|}
    \hline
    P    &         Итерации ($N=5$) & Ускорение ($N=5$) \\ \hline
	128 &        235.6   & 4.4      \\ \hline
	256 &       90.5    & 9.2      \\ \hline
	\end{tabular}

    \label{table:GKLS_RES_2}
\end{table}

\end{block}
          \begin{block}{Литература}
            \printbibliography

          \end{block}
        \end{column}
    \end{columns}
\end{frame}
\end{document}

### 2022-03-18 2022/Sensors/sensors.tex
add translated part

@@
 
-\cite{Strongin2000}
-Introduction
+Currently, machine learning (ML) methods are being used successfully to solve a wide range of problems in various application areas.
+One example of a class of problems where ML has demonstrated its effectiveness are the tasks of identifying the main properties of the phenomena, which are characterized by their stochastic nature or the presence of hidden parameters \cite{Golovenkin2020, Gonoskov2019, Kastalskiy2021SocialSD}.
+
+In many cases, solving application problems comes down to the problem of classification, i.e.
+assigning the object under study to one of the available classes.
+One of the well-known classification methods is the kNN (k-nearest neighbourhood) method.
+Although the earliest papers related to this method appeared more than $50$ years ago \cite{Cover1967, Cover1968}, theoretical studies of various aspects of using kNN are still going on [Gorban, Pestov].
+This is not to mention hundreds of publications in which this method is used to solve applied classification problems.
+Over many years of use, kNN has established itself as a simple and reliable method that yields acceptable results when solving many different problems.
+
+Among the undoubted merits of kNN is its "explainability", since the decision about whether a test object belongs to one class or another is clearly explained by similar properties of the test object and its nearest (in some metric) neighbours.
+Along with the study of the method’s theoretical properties, the issue of its effective implementation is also investigated.
+Thus, direct implementation of the method by using the  "brute-force" approach has computational complexity $O(n)$, where $n$ is the number of objects.
+Fast algorithms, which have computational complexity $O(log(n))$, use complex data structures of kd-tree type [kdtree-1,2] \cite{Bentley1975MultidimensionalBS}.
+
+One approach to accelerating the kNN algorithm is to use dimensionality reduction methods involving space-filling curves (Peano-Hilbert curves) \cite{Liao2001, Schubert2015FastAS}.
+In this paper, we develop this approach further and propose an algorithm that employs multiple space-filling curves and is faster (with comparable quality) compared to the algorithm using kd-trees.
+A specific method for constructing multiple Peano curves is outlined, and statements are given about the preservation of object proximity information in the course of dimensionality reduction.
+Comparison with known kNN implementations (in particular, those using kd-trees) was performed using test and real-life data with small dimensionality, i.e., just those data for which kd-trees show the best performance in terms of speed.
+
+The paper is organized as follows.
+Section 2 describes the scheme for constructing a Peano-Hilbert curve approximation (evolvent) and the kNN method, which uses the evolvent to reduce the dimensionality of the data.
+The fundamental drawbacks of this dimensionality reduction method are noted.
+Section 3 proposes a way to overcome the drawbacks mentioned, based on the use of multiple evolvents.
+A theoretical statement is formulated about preserving some proximity information in the multidimensional space on one of the one-dimensional scales.
+A kNN scheme using multiple evolvents is presented.
+Section 4 contains experimental results comparing different implementations of kNN using synthetic and real-life data.
+Section 5 concludes the paper.
 
@@
 
-\section{Methods}
+In this study, the classification problem will be understood as the problem of assigning an object $\omega$  to one of the predetermined classes based on its features.
+We will assume that each of the objects $\omega$ is represented as a feature vector
 
-\subsection{Core global search algorithm}\label{CoreGSA}
+\begin{equation}
+y = (y_1, y_2, ..., y_N) \in R^N,
+\end{equation}
+
+where the value of the $j$-th vector coordinate corresponds to the numerical value of the $j$-th feature of the object $\omega$.
+
+We will also assume that there is already some number of objects exactly classified (a test sample), i.e. for each object we know which class it belongs to.
+
+The $k$ nearest neighbours (kNN) method is based on the following simple rule: an object is considered to belong to the class to which most of its nearest neighbours belong.
+Here, "neighbours" refers to objects that are close (in one sense or another) to the one being studied.
+Note that in order to apply this method, some metric $L(\omega_i, \omega_j)$, i.e.
+the distance function, should be introduced in the feature space of objects.
+As a rule, the Euclidean distance is employed here, although other metrics may also be used [Gorban].
+
+The general scheme of object classification using kNN can be formulated as follows
+
+\begin{itemize}
+\item calculate the distance from the object being classified to each of the objects in the training set,
+\item select k objects of the training set with the minimum distance to them,
+\item the class of the object being classified is the class most often found among the k nearest neighbours.
+\end{itemize}
+
+In the practical implementation of kNN, an important indicator is the estimate of the time complexity of the neighbour finding procedure.
+The algorithms and data structures on which the search procedure is based include, in particular, the following.
+
+\begin{itemize}
+\item Brute force method.
+It is based on calculating all distances from the test object to other objects of the class and determining the smallest value.
+This method has complexity $O(n)$, where $n$ is the number of objects.
+\item KD-Tree method. It is based on a special kind of binary tree, where each node represents a point in the multidimensional space. The search procedure has complexity $O(log n)$.
+\item Ball-Tree method. It is yet another kind of tree structure, which has logarithmic complexity. This method is applicable to problems that have a large dimensionality.
+\end{itemize}
+
+The main drawback of the brute force method is the unacceptable running time and the rapid growth of the computation volume as the number of objects increases.
+This method can be applied when the number of objects is relatively small and the dimension of $y$ is small, and when the "curse of dimensionality" is not yet in full effect.
+
+For KD-Tree, the main drawback is that it slows down when the number of objects grows, which is caused by increasing complexity of the internal tree structure.
+However, KD-trees have proven to be a good solution for problems with small dimensionality, which will be considered in this study.
+
+As it was mentioned, the computational cost of searching for nearest objects increases with increasing dimensionality, both in the case of exhaustive search for all distances and when using tree structures.
+One mechanism for speeding up the search is dimensionality reduction, by reducing the problem of searching in a multidimensional space to the search in a one-dimensional space (over an interval).
+This is possible through the use of space-filling curves (Peano-Hilbert curves) that fill the multidimensional space.
+Such curves are used, for example, in multidimensional global optimization problems [ссылки], in image processing [ссылки], etc.
+
+An algorithm based on data curves can offer some advantages over tree-based algorithms by reducing the search time.
+Neighbour detection for test objects includes two phases: building and initialising the data structures (build step) and performing a search or query (query step).
+In most cases, the build phase occurs only once, while queries are repeated many times and can significantly affect the overall speed of the classification algorithm.
+
+Note that in practical classification tasks, different features can have different scales, which can significantly distort the real distance between objects.
+Therefore it will be assumed that prior to applying kNN data scaling has been performed
+
+\begin{equation}
+y_j \equiv \frac{y_j - y_{min}} {y_{max} - y_{min}} - \frac{1} {2}.
+\end{equation}
+
+Thus, the variation domain for all feature values will form a unit hypercube
+
+\begin{equation}
+D = \{ y \in R^N: -1/2 \leq y_j \leq 1/2,  1 \leq j \leq N \}.
+\end{equation}
+
+The $1/2$ hypercube offset  has been made for the convenience of further labelling.
+
+\section{Dimensionality reduction using space-filling curves}
+
+Let us briefly describe a general scheme for constructing a space-filling curve. We will consider here Hilbert's scheme for constructing such a curve.
+
+1. Divide hypercube $D$ with edge length of $1$ by coordinate hyperplanes into $2^N$ hypercubes of the first partition (with edge length of $1/2$).
+
+Then divide every hypercube of the first partition into $2^N$  hypercubes of the second partition (with edge length of $1/4$) by hyperplanes parallel to coordinate hyperplanes and passing through midpoints of hypercube edges orthogonal to them.
+
+By continuing the above process, i.e.
+by sequentially partitioning each subcube of the current partition into $2^N$ of the next partition, we can construct hypercubes of any $M$-th partition with edge length of $(1/2)^M$.
+The total number of subcubes of the $M$-th partition will be equal to $2^{NM}$.
+
+2.
+Now divide the segment $[0, 1]$ into $2^N$ equal parts, divide  each of them also into $2^N$ equal parts and so on.
+Denote the subinterval of the $M$-th partition by $d(M,v)$ , where $v$ is the coordinate of the left boundary point of the subinterval.
+Obviously, the length of the subinterval $d(M,v)$ will be equal to $2^{-NM}$.
+We will assume that the left boundary point belongs to the subinterval and the right boundary point does not.
+The only exception is the subinterval whose right boundary point is $1$, in which case it also belongs to the subinterval.
+
+3.
+Establish a one-to-one correspondence between subintervals and subcubes of the $M$-th partition.
+We denote by $D(M,v)$ a subcube corresponding to the subinterval $d(M,v)$.
+This correspondence should satisfy the following requirements.
+
+Condition 1. $D(M+1,v^\prime) \in D(M,v^{\prime\prime}) \;\: i.f.f. \;\:  d(M+1,v^\prime) \in d(M,v^{\prime\prime})$.
+
+Condition 2.
+Two subintervals $d(M,v^\prime)$ and $d(M,v^{\prime\prime})$ share a common endpoint i.f.f.
+the corresponding subcubes $D(M,v^\prime)$ and $D(M,v^{\prime\prime})$  are  contiguous, i.e.
+share a common edge.
+
+Condition 3.  If $x \in d(M,v)$ then $y(x) \in D(M,v), M \geq 1$.
+
+Note that the centre  $y^c(x)$ of the $M$-th partition subcube containing point $y(x)$ can be regarded as an approximation to $y(x)$ with the accuracy of $\varepsilon = 2^{-(M+1)}$, i.e.
+
+\begin{equation}
+\max_{1 \leq j \leq N} y_j^c(x)-y_j(x) \leq \varepsilon = 2^{-(M+1)}.
+\end{equation}
+
+The function $y^c(x)$ maps a uniform grid with step $2^{-NM}$, constructed in the interval $[0,1]$, onto a uniform grid with step $2^{-M}$, constructed in the hypercube $D$.
+A constructive way of establishing such a correspondence is described and theoretically justified in \cite{Strongin2000, Sergeyev2013}.
+
+Using the Peano-Hilbert curve, one can propose the following scheme for implementing the kNN algorithm.
+
+Let $\omega_i, 1 \leq i \leq S$, be the $i$-th object from the set of objects correlated with class $C_j$.
+Previously, a feature vector $y = (y_1, y_2, ..., y_N) \in R^N$, corresponding to each object under consideration was defined, i.e.
+
+\begin{equation}
+\omega_i \leftrightarrow y = (y_{i1}, y_{i2}, ..., y_{iN})
+\end{equation}
+
+Using the Peano curve, each object is assigned the value of $x_i$, i.e.
+
+\begin{equation}
+\omega_i \leftrightarrow x_i, \; y(x_i) = (y_{i1}, y_{i2}, ..., y_{iN}).
+\end{equation}
+
+Consider a test object $\omega^*$, to which the value $x^*$ is assigned.
+The degree of proximity of the test object $\omega^*$ to one of the objects $\omega_i$ is determined based on the distance on the one-dimensional scale $d_i = |x_i - x^* |$.
+Thus, the closest object to the test object has the following property:
+
+\begin{equation}
+d_{min} = \min \{ d_i: 1 \leq i \leq S \}.
+\end{equation}
+
+The operation of the algorithm to determine the nearest neighbour of the test object involves two phases: data preparation and the search itself.
+Data preparation consists of calculating the correspondence $\omega_i \rightarrow x_i$ for each input object, including the test object, and ordering the objects by the value of $x$.
+
+\begin{equation}
+x^1(\omega^1) \leq x^2(\omega^2) \leq ... \leq x^S(\omega^S).
+\end{equation}
+
+Since the corresponding object preimages $x_i \in [0,1]$ are ordered, the closest object can be found using a fast binary search algorithm.
+
+When $K$ neighbours need to be found, this search procedure can be repeated $K$ times.
+In this case it is necessary to exclude previously found items from the search.
+This can be done by excluding $x_i$ from the set being searched or by comparing the found $x_i$ with the previously found ones to find a match.
+For example, one possible implementation of searching for $K$ nearest neighbours could be as follows: in relation to $x^*$, we look for the nearest values to the left and right on the one-dimensional scale, choose the minimum one and continue searching sideways until exactly K neighbours are found.
+
+One obvious drawback of using the Peano curve in the kNN method is the loss of much of the information about the proximity of objects in multidimensional space when constructing their preimages on the one-dimensional scale.
+
+Indeed, the point $x \in [0,1]$ on the one-dimensional scale has only left and right neighbours, whereas its corresponding object $y(x) \in R^N$ can have neighbours along $N$ coordinate directions.
+As a result, when using Peano curve type mappings, close objects $y^\prime$, $y^{\prime\prime}$ in the $N$-dimensional space can have their quite distant corresponding preimages $x^\prime$, $x^{\prime\prime}$ on the interval $[0,1]$.
+In Fig.
+1 the green dots show two objects that are close in the two-dimensional space, while their preimages on the one-dimensional scale are far away from each other.
+The blue dots correspond to two objects, the distance between which is retained when changing to the one-dimensional scale.
+
+\begin{figure}
+\center
+\includegraphics[width=0.5\linewidth]{fig1.png}
+\caption{Peano curve}
+\label{fig1}
+\end{figure}
+
+This property makes dimensionality reduction using a single Peano curve practically inapplicable.
+To overcome this drawback, various approaches have been proposed, for example, the simultaneous use of two or more space-filling curves of different types [Scott1999], data shifting \cite{Liao2001}.
+In the following section, we propose a constructive way to preserve some information on object proximity during dimensionality reduction, based on the use of multiple same-type evolvents.
+
+\section{Constructing a family of Peano curves}
+
+Consider a family of Peano curves
+
+\begin{equation}
+Y_L(x)=\{ y^0(x), y^1(x), ..., y^L(x) \}
+\end{equation} 
+instead of a single curve $y(x)$. We will construct the family of curves as follows. Let us introduce a family of hypercubes
+
+\begin{equation}\label{hypercubes}
+D_i = \{ y \in R^N: -2^{-1} \leq y_i + 2^{-l} \leq 3 \cdot 2^{-1},  1 \leq i \leq N \}, \; 0 \leq l \leq L,
+\end{equation} 
+where the hypercube $D_{l+1}$ is obtained by "shifting" the hypercube $D_l$ along the main diagonal by step $2^{-l}$ for each coordinate.
+
+Let a Peano curve type mapping $y^0(x)$ map the interval $[0,1]$ onto the hypercube $D_0$ of (ссылка), i.e.
+
+\begin{equation}\label{hypercubeD0}
+D_0 = \{y^0(x): x \in [0, 1] \}.
+\end{equation}
+
+Any subcube of the $M$-th partition of the hypercube $D_0$ will have an edge of length $2^{-(M-1)}$ and will be denoted by $D_0(M, v)$, where $v$ is the left boundary point of the subinterval $d(M, v)$ corresponding to this subcube.
+Then the evolvents $y^l(x)=\{y_1^l(x), ..., y_N^l(x)\}$, whose coordinates are determined by the conditions
+
+\begin{equation}\label{maps}
+y_i^l(x) = y_i^{l-1}(x) + 2^{-l}, \; 1 \leq i \leq N, \; 1 \leq l \leq L,
+\end{equation} 
+map the interval $[0,1]$ onto the corresponding hypercubes  $D_l, \; 1 \leq l \leq L$ (the broken line in the figure shows the image of the interval $[0,1]$, which is obtained by using the evolvent $y^0(x), x \in [0,1]$).
+
+For any subcube $D_0(M,v)$ of the $M$-th partition of the hypercube $D_0$, there will exist a subcube $D_l(M,v)$ of the $M$-th partition of the hypercube $D_l$, and $D_l(M,v)$ can be obtained by shifting $D_0(M,v)$ along the main diagonal by the distance of $2^{-1}+2^{-2}+...+2^{-l}$.
+
+It follows from formulae (\ref{hypercubeD0}) and (\ref{maps}) that if an interval $d(M,v)$ is mapped onto the subcube $D_0(M,v)$, then there exists a family of subcubes
+
+\begin{equation}
+D_l(M,v) = y^l(d(M,v)), 1 \leq l \leq L,
+\end{equation} 
+connected with the corresponding subintervals $d_l(M,v_l) \subset [0,1]$, where $d(M,v) = d_0(M,v_0), v_0 = v$, such that
+
+\begin{equation}
+D_l(M,v) = y^l(d_l(M,v)), 1 \leq l \leq L,
+\end{equation}
+
+Since the hypercube $D$ of (ссылка) belongs to the general part of the family of hypercubes (\ref{hypercubes}) (the boundary of the hypercube $D$ is highlighted in Fig \ref{fig2}), then, by introducing an additional function
+
+\begin{equation}
+g_0(y) = \max \{ |y_i|-2^{-1}: \; 1 \leq i \leq N \},
+\end{equation} 
+the original hypercube $D$ can be represented as
+
+\begin{equation}
+D = \{y^l(x): x \in [0,1], \; g_0(y^l(x)) \leq 0 \}, \; 0 \leq l \leq L,
+\end{equation} 
+i.e. $g_0(y) \leq 0$, if $y \in D$, and $g_0(y) > 0$ if otherwise. Hence, any point $y \in D$ has its preimage $x^l \in [0,1]$ at every mapping $y^l(x), \; 0 \leq l \leq L$, i.e.
+
+\begin{equation}
+y = y^l(x^l), \; x^l \in [0, 1], \; 0 \leq l \leq L.
+\end{equation}
+
+By using multiple mappings $y^l(x), \; 0 \leq l \leq L$, the following relationship between neighborhoods on one-dimensional scales and neighborhoods in the original multidimensional domain is defined.
+
+\textbf{Theorem 1.} Let $y^*$ be an arbitrary point from the domain $D$ belonging to the interval with endpoints $y^\prime, y^{\prime\prime} \in D$, which differ in the values of the only coordinate, and let
+
+\begin{equation}
+|y_j^\prime - y^{\prime\prime}_j| \leq 2^{-p}, \; y^\prime_i = y^{\prime\prime}_i = y_i^*, \; 1 \leq i \leq N, i \ne j,
+\end{equation} 
+where $p$ is an integer, $1 \leq p \leq L-1$, and $j$ is the number of coordinate whose values for points $y^*, y^\prime, y^{\prime\prime}$ are different.
+Then there exists at least one correspondence $y^l(x), 0 \leq l \leq L$, and preimages $x^*, x^\prime, x^{\prime\prime} \in [0, 1]$ such that
+
+\begin{equation}
+\begin{array}
+\: y^* = y^l(x^*), y^\prime = y^l(x^\prime), y^{\prime\prime} = y^l(x^{\prime\prime}), \\
+\max \{ |x^\prime - x^*|, |x^{\prime\prime} - x^*|, |x^\prime - x^{\prime\prime}| \} \leq 2^{-pN}.
+\end{array}
+\end{equation}
+
+\textbf{Remark}.
+The conditions of the theorem distinguish a specific neighborhood of the point $y^*$.
+This neighborhood comprises only such neighbours of this point which can be obtained by  shifting $y^*$ parallel to one of the coordinate axes by a distance not exceeding $2^{-p}$.
+By varying the value of $j, 1 \leq j \leq N$, under the conditions of the theorem, one can identify the nearest neighbours of point $y^*$ along $N$ coordinate directions.
+According to the statement, the proximity of points in the $N$-dimensional space in a particular direction will be reflected by the proximity of their preimages on one of the one-dimensional segments.
+In this case, the corresponding one-dimensional intervals can be different for different directions.
+
+The proof of this theorem is given in \cite{Strongin2000}.
+
+As an illustration, Figure \ref{fig2} shows a family of space-filling curves in extended domains $D_l$, where a square with a dark border highlights the hypercube $D$ belonging to each extended domain.
+The dots indicate proximate objects in the domain $D$, the distance between which will vary depending on the one-dimensional scale used.
+The effect of maintaining distance proximity on one of the scales is clearly visible.
+For example, objects marked with green dots will be far away on the first two one-dimensional scales and close to each other on the third one.
+Objects marked in blue will be far away on the first and third scales and close to each other on the second scale.
+
+\begin{figure}	
+\widefigure
+\includegraphics[width=1.0\linewidth]{fig2.png}
+\caption{Family of Peano curves\label{fig2}}
+\end{figure}
+
+The scheme of implementation of the kNN algorithm using multiple Peano curves will look as follows.
+
+For every object $\omega_i$ a set of its preimages on different one-dimensional scales is assigned
+
+\begin{equation}
+\omega_i \rightarrow \{ x_i^0, x_i^1, ..., x_i^L \}
+\end{equation} 
+where $l, \; 0 \leq l \leq L$, is the number of the one-dimensional interval.
+
+The degree of proximity of a test object to one of the known objects can be expressed in terms of the distance on the $l$-th scale.
+
+\begin{equation}
+d_i^n = |x_i^n - x^{*n}|
+\end{equation}
+
+Thus, the distance $d(\omega_i,\omega_j)$ between two objects $\omega_i$ and $\omega_j$ on a one-dimensional scale will correspond to the minimum distance among all the mappings.
+If we find the minimum distance among all these distances, we get the nearest neighbour
+
+\begin{equation}
+d_{min} = \min \{ d_i^n \} = \min \{ |x_i^n - x^{*n}| \}.
+\end{equation}
+
+Next, the search procedure needs to be run as many times as there are nearest neighbours to be identified.
+The main problem is that an object previously found on one scale can be found again, including on another scale.
+This requires either removing the values of x found objects from all scales or searching for $K$ nearest neighbours within each scale and then merging the results.
+
+\section{Special features of the implementation of the kNN method using multiple space-filling curves}
+
+The procedure of finding $K$ nearest neighbours using multiple space-filling curves was implemented in two ways.
+The first of them (denoted hereafter as Algorithm A) involves searching for $K$ neighbours on each scale and then merging the results and selecting $K$ neighbours from all scales.
+The peculiarity of this algorithm is that there is no procedure to remove the objects found from the scales.
+
+The second way (Algorithm B) involves finding the nearest object to the test object on each scale, determining the nearest object among those found, and then removing the selected object from all scales.
+In this case, the search for the next nearest neighbour is performed by repeatedly calling the same procedure.
+After all nearest neighbours are determined, the previously excluded objects are returned to their places and the algorithm is ready to work with the next test object.
+
+Here is a more detailed description of these algorithms.
+\\[12pt]
+\textit{Data preparation}
+\\[12pt]
+To run Algorithms A and B, we need to compute the preimages $\{x_i^0, x_i^1, ..., x_i^L\}$ on one-dimensional scales for all objects $\omega_i$, and order them in ascending order.
+It is convenient to use RB-tree-based data structures to store ordered one-dimensional values.
+The number of instances of tree structures is equal to the number of scales.
+
+The input data for kNN are the values of object features imported into the program from external files. A text label indicating the class to which the object belongs is read with each such object.
+
+The input data of the algorithm also includes the feature vector $y^*$ corresponding to the test object $\omega^*$.
+\\[12pt]
+\underline{Procedure A}
+
+\begin{itemize}
+\item For each scale, search for $K$ objects closest to $\omega^*$. The result of the search is a set of vectors of  the nearest objects on each scale.
+\item The vectors of the objects found are inserted into an RB-tree according to the distance criterion $d_{min}$.
+\item After processing all the scales, the first $K$ objects are extracted from the resulting tree and they are considered to be the search result.
+\item The number of objects belonging to each class is counted and the decision is made whether   the test object belongs to a particular class.
+\end{itemize}
+\\[12pt]
+\underline{Procedure B}
+
+\begin{itemize}
+\item For each scale, search for the closest object to $ω^*$; the result is a set of distances between the test object and the closest object for each scale.
+\item The object with minimal distance on all scales is selected.
+\item This object is saved in a special vector and excluded from the set of objects.
+\item The whole procedure is repeated $K$ times, resulting in a vector of K nearest neighbours.
+\item The number of objects belonging to each class is counted and the decision is made whether   the test object belongs to a particular class.
+\item If re-classification is required, the procedure is called to recover the previously excluded objects and search for neighbours.
+\end{itemize}
 
@@
 
-\begin{figure}
-\center
-\includegraphics[width=0.5\linewidth]{fig1.png}
-\caption{Using GSA-DT algorithm to solve GKLS problems}
-\label{fig1}
-\end{figure}
-
-some text
-
- Example of a figure that spans the whole page width. The same concept works for tables, too.
-
 \end{paracol}
-
-\begin{figure}	
-\widefigure
-\includegraphics[width=1.0\linewidth]{fig2.png}
-\caption{This is a wide figure.\label{fig2}}
-\end{figure}
 

### 2022-03-19 2022/Sensors/sensors.tex
fix some mistakes and references

@@
 Currently, machine learning (ML) methods are being used successfully to solve a wide range of problems in various application areas.
-One example of a class of problems where ML has demonstrated its effectiveness are the tasks of identifying the main properties of the phenomena, which are characterized by their stochastic nature or the presence of hidden parameters \cite{Golovenkin2020, Gonoskov2019, Kastalskiy2021SocialSD}.
+One example of a class of problems where ML has demonstrated its effectiveness are the tasks of identifying the main properties of the phenomena, which are characterized by their stochastic nature or the presence of hidden parameters \cite{Golovenkin2020, Gonoskov2019, Kastalskiy2021}.
 
@@
 
-Among the undoubted merits of kNN is its "explainability", since the decision about whether a test object belongs to one class or another is clearly explained by similar properties of the test object and its nearest (in some metric) neighbours.
-Along with the study of the method’s theoretical properties, the issue of its effective implementation is also investigated.
-Thus, direct implementation of the method by using the  "brute-force" approach has computational complexity $O(n)$, where $n$ is the number of objects.
+Among the undoubted merits of kNN is its ``explainability'', since the decision about whether a test object belongs to one class or another is clearly explained by similar properties of the test object and its nearest (in some metric) neighbours.
+Along with the study of the method's theoretical properties, the issue of its effective implementation is also investigated.
+Thus, direct implementation of the method by using the  ``brute-force'' approach has computational complexity $O(n)$, where $n$ is the number of objects.
 Fast algorithms, which have computational complexity $O(log(n))$, use complex data structures of kd-tree type [kdtree-1,2] \cite{Bentley1975MultidimensionalBS}.
@@
 We will assume that each of the objects $\omega$ is represented as a feature vector
-
 \begin{equation}
@@
 \end{equation} 
-
 where the value of the $j$-th vector coordinate corresponds to the numerical value of the $j$-th feature of the object $\omega$.
@@
 
-The general scheme of object classification using kNN can be formulated as follows
-
+The general scheme of object classification using kNN can be formulated as follows:
 \begin{itemize}
-\item calculate the distance from the object being classified to each of the objects in the training set,
-\item select k objects of the training set with the minimum distance to them,
-\item the class of the object being classified is the class most often found among the k nearest neighbours.
+\item calculate the distance from the object being classified to each of the objects in the training set;
+\item select k objects of the training set with the minimum distance to them;
+\item the class of the object being classified is the class most often found among the $k$ nearest neighbours.
 \end{itemize}
@@
 The main drawback of the brute force method is the unacceptable running time and the rapid growth of the computation volume as the number of objects increases.
-This method can be applied when the number of objects is relatively small and the dimension of $y$ is small, and when the "curse of dimensionality" is not yet in full effect.
+This method can be applied when the number of objects is relatively small and the dimension of $y$ is small, and when the ``curse of dimensionality'' is not yet in full effect.
 
@@
 This is possible through the use of space-filling curves (Peano-Hilbert curves) that fill the multidimensional space.
-Such curves are used, for example, in multidimensional global optimization problems [ссылки], in image processing [ссылки], etc.
+Such curves are used, for example, in multidimensional global optimization problems \cite{Strongin2018,Gergel2021}, in image processing \cite{Liang2008,Costa2012,Herrero2015}, etc.
 
@@
 Therefore it will be assumed that prior to applying kNN data scaling has been performed
-
 \begin{equation}
@@
 Thus, the variation domain for all feature values will form a unit hypercube
-
-\begin{equation}
+\begin{equation}\label{D}
 D = \{ y \in R^N: -1/2 \leq y_j \leq 1/2,  1 \leq j \leq N \}.
 \end{equation} 
-
 The $1/2$ hypercube offset  has been made for the convenience of further labelling.
@@
 Note that the centre  $y^c(x)$ of the $M$-th partition subcube containing point $y(x)$ can be regarded as an approximation to $y(x)$ with the accuracy of $\varepsilon = 2^{-(M+1)}$, i.e.
-
 \begin{equation}
@@
 Previously, a feature vector $y = (y_1, y_2, ..., y_N) \in R^N$, corresponding to each object under consideration was defined, i.e.
-
 \begin{equation}
@@
 Using the Peano curve, each object is assigned the value of $x_i$, i.e.
-
 \begin{equation}
@@
 Thus, the closest object to the test object has the following property:
-
 \begin{equation}
@@
 Data preparation consists of calculating the correspondence $\omega_i \rightarrow x_i$ for each input object, including the test object, and ordering the objects by the value of $x$.
-
 \begin{equation}
@@
 This property makes dimensionality reduction using a single Peano curve practically inapplicable.
-To overcome this drawback, various approaches have been proposed, for example, the simultaneous use of two or more space-filling curves of different types [Scott1999], data shifting \cite{Liao2001}.
+To overcome this drawback, various approaches have been proposed, for example, the simultaneous use of two or more space-filling curves of different types \cite{Scott1999}, data shifting \cite{Liao2001}.
 In the following section, we propose a constructive way to preserve some information on object proximity during dimensionality reduction, based on the use of multiple same-type evolvents.
@@
 Consider a family of Peano curves
-
 \begin{equation}
@@
 instead of a single curve $y(x)$. We will construct the family of curves as follows. Let us introduce a family of hypercubes
-
 \begin{equation}\label{hypercubes}
@@
 
-Let a Peano curve type mapping $y^0(x)$ map the interval $[0,1]$ onto the hypercube $D_0$ of (ссылка), i.e.
-
+Let a Peano curve type mapping $y^0(x)$ map the interval $[0,1]$ onto the hypercube $D_0$ of (\ref{hypercubes}), i.e.
 \begin{equation}\label{hypercubeD0}
@@
 Then the evolvents $y^l(x)=\{y_1^l(x), ..., y_N^l(x)\}$, whose coordinates are determined by the conditions
-
 \begin{equation}\label{maps}
@@
 It follows from formulae (\ref{hypercubeD0}) and (\ref{maps}) that if an interval $d(M,v)$ is mapped onto the subcube $D_0(M,v)$, then there exists a family of subcubes
-
 \begin{equation}
@@
 connected with the corresponding subintervals $d_l(M,v_l) \subset [0,1]$, where $d(M,v) = d_0(M,v_0), v_0 = v$, such that 
-
 \begin{equation}
@@
 
-Since the hypercube $D$ of (ссылка) belongs to the general part of the family of hypercubes (\ref{hypercubes}) (the boundary of the hypercube $D$ is highlighted in Fig \ref{fig2}), then, by introducing an additional function
-
+Since the hypercube $D$ from (\ref{D}) belongs to the common part of the family of hypercubes (\ref{hypercubes}) (the boundary of the hypercube $D$ is highlighted in Fig.~\ref{fig2}), then, by introducing an additional function
 \begin{equation}
@@
 the original hypercube $D$ can be represented as
-
 \begin{equation}
@@
 i.e. $g_0(y) \leq 0$, if $y \in D$, and $g_0(y) > 0$ if otherwise. Hence, any point $y \in D$ has its preimage $x^l \in [0,1]$ at every mapping $y^l(x), \; 0 \leq l \leq L$, i.e.
-
 \begin{equation}
@@
 \begin{equation}
-\begin{array}
-\: y^* = y^l(x^*), y^\prime = y^l(x^\prime), y^{\prime\prime} = y^l(x^{\prime\prime}), \\
+ y^* = y^l(x^*), y^\prime = y^l(x^\prime), y^{\prime\prime} = y^l(x^{\prime\prime}), \\
 \max \{ |x^\prime - x^*|, |x^{\prime\prime} - x^*|, |x^\prime - x^{\prime\prime}| \} \leq 2^{-pN}.
-\end{array}
 \end{equation}
@@
 
+\end{paracol}
+
 \begin{figure}	
@@
 \end{figure}
+
+\begin{paracol}{2}
+\linenumbers
+\switchcolumn
 
@@
 The input data of the algorithm also includes the feature vector $y^*$ corresponding to the test object $\omega^*$.
-\\[12pt]
+
+%\\[12pt]
 \underline{Procedure A}
@@
 \end{itemize}
-\\[12pt]
+
+%\\[12pt]
 \underline{Procedure B}
@@
 \begin{itemize}
-\item For each scale, search for the closest object to $ω^*$; the result is a set of distances between the test object and the closest object for each scale.
+\item For each scale, search for the closest object to $\omega^*$; the result is a set of distances between the test object and the closest object for each scale.
 \item The object with minimal distance on all scales is selected.

### 2022-03-20 2022/Sensors/sensors.tex
add two references

@@
 Currently, machine learning (ML) methods are being used successfully to solve a wide range of problems in various application areas.
-One example of a class of problems where ML has demonstrated its effectiveness are the tasks of identifying the main properties of the phenomena, which are characterized by their stochastic nature or the presence of hidden parameters \cite{Golovenkin2020, Gonoskov2019, Kastalskiy2021}.
+One example of a class of problems where ML has demonstrated its effectiveness are the tasks of identifying the main properties of the phenomena, which are characterized by their stochastic nature or the presence of hidden parameters \cite{Golovenkin2020,Gonoskov2019,Kastalskiy2021}.
 
@@
 One of the well-known classification methods is the kNN (k-nearest neighbourhood) method.
-Although the earliest papers related to this method appeared more than $50$ years ago \cite{Cover1967, Cover1968}, theoretical studies of various aspects of using kNN are still going on [Gorban, Pestov].
+Although the earliest papers related to this method appeared more than $50$ years ago \cite{Cover1967, Cover1968}, theoretical studies of various aspects of using kNN are still going on \cite{Pestov2013,Mirkes2020}.
 This is not to mention hundreds of publications in which this method is used to solve applied classification problems.
@@
 Thus, direct implementation of the method by using the  ``brute-force'' approach has computational complexity $O(n)$, where $n$ is the number of objects.
-Fast algorithms, which have computational complexity $O(log(n))$, use complex data structures of kd-tree type [kdtree-1,2] \cite{Bentley1975MultidimensionalBS}.
+Fast algorithms, which have computational complexity $O(log(n))$, use complex data structures of kd-tree type \cite{Bentley1975MultidimensionalBS}.
 
@@
 
-By using multiple mappings $y^l(x), \; 0 \leq l \leq L$, the following relationship between neighborhoods on one-dimensional scales and neighborhoods in the original multidimensional domain is defined.
+By using multiple mappings $y^l(x), \; 0 \leq l \leq L$, the following relationship between neighbourhoods on one-dimensional scales and neighbourhoods in the original multidimensional domain is defined.
 
 \textbf{Theorem 1.} Let $y^*$ be an arbitrary point from the domain $D$ belonging to the interval with endpoints $y^\prime, y^{\prime\prime} \in D$, which differ in the values of the only coordinate, and let
-
 \begin{equation}
@@
 Then there exists at least one correspondence $y^l(x), 0 \leq l \leq L$, and preimages $x^*, x^\prime, x^{\prime\prime} \in [0, 1]$ such that
-
 \begin{equation}
@@
 For every object $\omega_i$ a set of its preimages on different one-dimensional scales is assigned 
-
 \begin{equation}
@@
 The degree of proximity of a test object to one of the known objects can be expressed in terms of the distance on the $l$-th scale.
-
 \begin{equation}
@@
 If we find the minimum distance among all these distances, we get the nearest neighbour
-
 \begin{equation}
@@
 
-The input data of the algorithm also includes the feature vector $y^*$ corresponding to the test object $\omega^*$.
+The input data of the algorithm also includes the feature vector $y^*$ corresponding to the test object $\omega^*$.\\
 
@@
 
-\end{paracol}
-
-\begin{paracol}{2}
-\linenumbers
-\switchcolumn
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Conceptualization and methodology, K.B.; software and validation, A.S., A.Sh.; formal analysis, K.B.; investigation, A.S.; data curation, A.Sh.; writing --- original draft preparation, K.B.; writing- -- review and editing, K.B.; visualization, A.S.; funding acquisition, K.B.
+\authorcontributions{Conceptualization and methodology, K.B.; software and validation, A.S., A.Sh.; formal analysis, K.B.; investigation, A.S.; data curation, A.Sh.; writing --- original draft preparation, K.B.; writing --- review and editing, K.B.; visualization, A.S.; funding acquisition, K.B.
 All authors have read and agreed to the published version of the manuscript.}

### 2022-03-21 2022/Sensors/sensors.tex
two new references

@@
 %=================================================================
-\documentclass[sensors,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+\documentclass[entropy,article,submit,moreauthors,pdftex]{Definitions/mdpi}
 
@@
 the distance function, should be introduced in the feature space of objects.
-As a rule, the Euclidean distance is employed here, although other metrics may also be used [Gorban].
+As a rule, the Euclidean distance is employed here, although other metrics may also be used \cite{Mirkes2020}.
 
@@
 This is possible through the use of space-filling curves (Peano-Hilbert curves) that fill the multidimensional space.
-Such curves are used, for example, in multidimensional global optimization problems \cite{Strongin2018,Gergel2021}, in image processing \cite{Liang2008,Costa2012,Herrero2015}, etc.
+Such curves are used, for example, in global optimization \cite{Lera2015,Lera2018,Strongin2018,Gergel2021}, in numerical approximation of solutions to systems of nonlinear inequalities \cite{Lera2021}, in image processing \cite{Liang2008,Costa2012,Herrero2015}, etc.
 

### 2022-03-21 2022/PAVTOneAPI/PAVT_Poster/poster.tex
Обновленный плакат.

@@
     \begin{columns}[t]
-        \begin{column}[t]{0.48\paperwidth}
+        \begin{column}[t]{0.47\paperwidth}
             \begin{block}{Постановка задачи}
              Задача многомерной многоэкстремальной оптимизации может быть определена как поиск наименьшего значения действительной функции \(\phi(y)\)  в гиперинтервале \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
-Будем предполагать, что целевая функция задана как <<черный ящик>> и удовлетворяет условию Липшица с априори неизвестной константой \(L\)
+
+             Будем предполагать, что целевая функция задана по принципу <<черного ящика>> и удовлетворяет условию Липшица с априори неизвестной константой \(L\).
 
@@
 
- \begin{minipage}[t]{.48\textwidth}
+ \begin{minipage}[t]{.47\textwidth}
               \begin{figure}
                   \centering
-                  \includegraphics[scale=1.27]{images/pean_2.jpg}
+                  \includegraphics[scale=1.49]{images/pean_2.jpg}
               \end{figure}
@@
 
+Распараллеливание организовано следующим образом: в ходе выполнения одной итерации алгоритма одновременно проводятся \(P \geq 1\) испытаний в различных точках области поиска.
+Важно отметить, что рассматриваются функции, в которых самым трудоемким и времязатратным этапом работы является это вычисление значения функции в точке.
+Поэтому организация одновременных вычислений этих значений положительно влияет на скорость работы алгоритма для таких функций.
+
+ \begin{minipage}[t]{.47\textwidth}
+              \begin{figure}
+                  \centering
+                  \includegraphics[scale=2.05]{images/PlotGS_1.png}
+              \end{figure}
+              \end{minipage}
+\end{block}
+
+        \end{column}
+        \begin{column}[t]{0.47\paperwidth}
+          \begin{block}{Общая схема алгоритма}
+
 Общая схема одной итерации одномерного метода:
@@
                 \item Вычислить для каждого интервала \((x_{i-1};x_{i}),1\leq i\leq k\)  характеристику \(R(i)\) .
-                \item Определить интервал \((x_{t-1};x_{t})\) , которому соответствует максимальная характеристика \(R(t)=\max\{R(i),1\leq i\leq k\}\).
-                \item Провести следующее испытание в точке \(x^{k+1}=d(t)\in (x_{t-1};x_{t})\) , где \(d(t)\)  — правило размещения точки следующего испытания в интервале с номером \(t\).
+                \item Упорядочить характеристики \(R(i)\) в порядке убывания и выбрать \(P \) наибольших из них с номерами интервалов \(t_{j}, 1\leqslant j \leqslant P: R(t_{1})\geqslant R(t_{2})\geqslant ...
+\geqslant  R(t_{k-1})\geqslant R(t_{k})\)
+                \item Параллельно провести следующие испытания в точках \(x^{k+j}, 1\leqslant j \leqslant P, \) \(x^{k+j}=d(t)\in (x_{t-1};x_{t})\) , где \(d(t)\)  — правило размещения точки следующего испытания в интервале с номером \(t\).
                 \item Проверить выполнение критерия остановки \(x_{t}-x_{t-1}<\varepsilon\).
               \end{enumerate}
-
- \begin{minipage}[t]{.48\textwidth}
-              \begin{figure}
-                  \centering
-                  \includegraphics[scale=1.27]{images/PlotGS.png}
-              \end{figure}
-              \end{minipage}
-\end{block}
-
-        \end{column}
-        \begin{column}[t]{0.48\paperwidth}
-          \begin{block}{Организация распараллеливания}
-
-Распараллеливание организовано следующим образом: в ходе выполнения одной итерации алгоритма одновременно проводятся \(P \geq 1\) испытаний в различных точках области поиска.
-Важно отметить, что рассматриваются функции, в которых самым трудоемким и времязатратным этапом работы является это вычисление значения функции в точке.
-Поэтому организация одновременных вычислений этих значений положительно влияет на скорость работы алгоритма для таких функций.
 
@@
 На данный момент получены результаты вычислительных экспериментов, в которых тестовые задачи формируются специальным генератором \cite{GKLS}.
-В таблицах \ref{table:GKLS_RES_1}, \ref{table:GKLS_RES_2} приведены среднее число итераций параллельного алгоритма глобального поиска и ускорение относительно последовательной версии.
+В таблицах \ref{table:GKLS_RES_1}, \ref{table:GKLS_RES_1} приведены среднее число итераций параллельного алгоритма глобального поиска и ускорение относительно последовательной версии.
 В каждом эксперименте решалось 100 задач размерности \(N = 4\) и \(N = 5\).
 Эксперименты проведены на компьютере с процессором Intel Core I5-7300HQ 2.5 GHz с видеокартой HD Graphics 630, 16 GB RAM, использовался язык \textit{data parallel c++}.
+
+%{\setlength{\extrarowheight}{5pt}
 
@@
     \caption{Число итераций и ускорение при решении серии задач размерности $N=4$}
-    \begin{tabular}{|c|c|c|}
+     \renewcommand{\arraystretch}{1.4}
+    \renewcommand{\tabcolsep}{1cm}
+    \begin{tabular}{|c|c|c|c|}
     \hline
-    P    & Итерации ($N=4$) & Ускорение ($N=4$) \\ \hline
-	128 & 106.5   & 4.7    \\ \hline
-	256 & 41.3    & 8.5          \\ \hline
+    P  & Итерации  & Время работы (сек.) & Ускорение \\ \hline
+	128 & 106.5 & 1.2  & 4.7    \\ \hline
+	256 & 41.3 & 0.66   & 8.5          \\ \hline
 	\end{tabular}
@@
     \caption{Число итераций и ускорение при решении серии задач размерности $N=5$}
-    \begin{tabular}{|c|c|c|}
+    \renewcommand{\arraystretch}{1.4}
+    \renewcommand{\tabcolsep}{1cm}
+    \begin{tabular}{|c|c|c|c|}
     \hline
-    P    &         Итерации ($N=5$) & Ускорение ($N=5$) \\ \hline
-	128 &        235.6   & 4.4      \\ \hline
-	256 &       90.5    & 9.2      \\ \hline
+    P  & Итерации  & Время работы (сек.) & Ускорение \\ \hline
+	128 &        235.6 & 2.6  & 4.4      \\ \hline
+	256 &       90.5  & 1.2 & 9.2      \\ \hline
 	\end{tabular}

### 2022-03-22 2022/Sensors/sensors.tex
fix some typos

@@
 % Keywords
-\keyword{space-filling curves}
+\keyword{machine learning, kNN, dimensionality reduction, multiple space-filling curves}
 
@@
 This method has complexity $O(n)$, where $n$ is the number of objects.
-\item KD-Tree method. It is based on a special kind of binary tree, where each node represents a point in the multidimensional space. The search procedure has complexity $O(log n)$.
-\item Ball-Tree method. It is yet another kind of tree structure, which has logarithmic complexity. This method is applicable to problems that have a large dimensionality.
+\item KD-tree method. It is based on a special kind of binary tree, where each node represents a point in the multidimensional space. The search procedure has complexity $O(log(n))$.
+\item Ball-tree method. It is yet another kind of tree structure, which has logarithmic complexity. This method is applicable to problems that have a large dimensionality.
 \end{itemize}
@@
 
-For KD-Tree, the main drawback is that it slows down when the number of objects grows, which is caused by increasing complexity of the internal tree structure.
-However, KD-trees have proven to be a good solution for problems with small dimensionality, which will be considered in this study.
+For kd-tree, the main drawback is that it slows down when the number of objects grows, which is caused by increasing complexity of the internal tree structure.
+However, kd-trees have proven to be a good solution for problems with small dimensionality, which will be considered in this study.
 
@@
 \begin{equation}
-y_j \equiv \frac{y_j - y_{min}} {y_{max} - y_{min}} - \frac{1} {2}.
+y_j = \frac{y_j - y_{min}} {y_{max} - y_{min}} - \frac{1} {2}.
 \end{equation}
@@
 
-1. Divide hypercube $D$ with edge length of $1$ by coordinate hyperplanes into $2^N$ hypercubes of the first partition (with edge length of $1/2$).
+1. Divide hypercube $D$ form (\ref{D}) with edge length of $1$ by coordinate hyperplanes into $2^N$ hypercubes of the first partition (with edge length of $1/2$).
 
@@
 Now divide the segment $[0, 1]$ into $2^N$ equal parts, divide  each of them also into $2^N$ equal parts and so on.
-Denote the subinterval of the $M$-th partition by $d(M,v)$ , where $v$ is the coordinate of the left boundary point of the subinterval.
+Denote the subinterval of the $M$-th partition by $d(M,v)$, where $v$ is the coordinate of the left boundary point of the subinterval.
 Obviously, the length of the subinterval $d(M,v)$ will be equal to $2^{-NM}$.
@@
 
-Condition 1. $D(M+1,v^\prime) \in D(M,v^{\prime\prime}) \;\: i.f.f. \;\:  d(M+1,v^\prime) \in d(M,v^{\prime\prime})$.
+Condition 1. $D(M+1,v^\prime) \in D(M,v^{\prime\prime})$ i.f.f. $d(M+1,v^\prime) \in d(M,v^{\prime\prime})$.
 
@@
 \begin{equation}
-\max_{1 \leq j \leq N} y_j^c(x)-y_j(x) \leq \varepsilon = 2^{-(M+1)}.
+\max_{1 \leq j \leq N} \left| y_j^c(x)-y_j(x) \right| \leq \varepsilon = 2^{-(M+1)}.
 \end{equation}
@@
 \end{equation} 
-where the hypercube $D_{l+1}$ is obtained by "shifting" the hypercube $D_l$ along the main diagonal by step $2^{-l}$ for each coordinate.
+where the hypercube $D_{l+1}$ is obtained by shifting the hypercube $D_l$ along the main diagonal by step $2^{-l}$ for each coordinate.
 
@@
 \end{equation} 
-map the interval $[0,1]$ onto the corresponding hypercubes  $D_l, \; 1 \leq l \leq L$ (the broken line in the figure shows the image of the interval $[0,1]$, which is obtained by using the evolvent $y^0(x), x \in [0,1]$).
+map the interval $[0,1]$ onto the corresponding hypercubes  $D_l, \; 1 \leq l \leq L$ (the broken line in Fig.~\ref{fig2} shows the image of the interval $[0,1]$, which is obtained by using the evolvent $y^0(x), x \in [0,1]$).
 
-For any subcube $D_0(M,v)$ of the $M$-th partition of the hypercube $D_0$, there will exist a subcube $D_l(M,v)$ of the $M$-th partition of the hypercube $D_l$, and $D_l(M,v)$ can be obtained by shifting $D_0(M,v)$ along the main diagonal by the distance of $2^{-1}+2^{-2}+...+2^{-l}$.
+For any subcube $D_0(M,v)$ of the $M$-th partition of the hypercube $D_0$, there will exist a subcube $D_l(M,v)$ of the $M$-th partition of the hypercube $D_l$, and $D_l(M,v)$ can be obtained by shifting $D_0(M,v)$ along the main diagonal by the distance of
+\[
+2^{-1}+2^{-2}+...+2^{-l}.
+\]
 
@@
 \begin{equation}
-D_l(M,v) = y^l(d_l(M,v)), 1 \leq l \leq L,
+D_l(M,v) = y^l(d_l(M,v)), 1 \leq l \leq L.
 \end{equation}
@@
 
-As an illustration, Figure \ref{fig2} shows a family of space-filling curves in extended domains $D_l$, where a square with a dark border highlights the hypercube $D$ belonging to each extended domain.
+As an illustration, Fig.
+\ref{fig2} shows a family of space-filling curves in extended domains $D_l$, where a square with a dark border highlights the hypercube $D$ belonging to each extended domain.
 The dots indicate proximate objects in the domain $D$, the distance between which will vary depending on the one-dimensional scale used.
@@
 
-The degree of proximity of a test object to one of the known objects can be expressed in terms of the distance on the $l$-th scale.
+The degree of proximity of a test object to one of the known objects can be expressed in terms of the distance
 \begin{equation}
-d_i^n = |x_i^n - x^{*n}|
+d_i^l = |x_i^l - x^{*l}|
 \end{equation} 
+on the $l$-th scale.
 
@@
 \begin{equation}
-d_{min} = \min \{ d_i^n \} = \min \{ |x_i^n - x^{*n}| \}.
+d_{min} = \min \{ d_i^l \} = \min \{ |x_i^l - x^{*l}| \}.
 \end{equation}
@@
 The main problem is that an object previously found on one scale can be found again, including on another scale.
-This requires either removing the values of x found objects from all scales or searching for $K$ nearest neighbours within each scale and then merging the results.
+This requires either removing the values of $x$ found objects from all scales or searching for $K$ nearest neighbours within each scale and then merging the results.
 

### 2022-03-23 2022/Sensors/sensors.tex
fix one more mistake

@@
 
-The degree of proximity of a test object to one of the known objects can be expressed in terms of the distance
+The degree of proximity of a test object $\omega^*$ to one of the known objects $\omega_i$ can be expressed in terms of the distance
 \begin{equation}
@@
 
-Thus, the distance $d(\omega_i,\omega_j)$ between two objects $\omega_i$ and $\omega_j$ on a one-dimensional scale will correspond to the minimum distance among all the mappings.
-If we find the minimum distance among all these distances, we get the nearest neighbour
+Thus, the distance $d(\omega_i,\omega^*)$ between two objects $\omega_i$ and $\omega^*$ on a one-dimensional scale will correspond to the minimum distance among all the mappings 
+\[
+d(\omega_i,\omega^*) = \min_{0\leq l\leq L} d_i^l = \min_{0\leq l\leq L}  |x_i^l - x^{*l}| .
+\]
+If we find the minimum distance
 \begin{equation}
-d_{min} = \min \{ d_i^l \} = \min \{ |x_i^l - x^{*l}| \}.
+d_{min} = \min_{1\leq i\leq n} d(\omega_i,\omega^*) ,
 \end{equation} 
+we get the nearest neighbour.
 
@@
 \item This object is saved in a special vector and excluded from the set of objects.
-\item The whole procedure is repeated $K$ times, resulting in a vector of K nearest neighbours.
+\item The whole procedure is repeated $K$ times, resulting in a vector of $K$ nearest neighbours.
 \item The number of objects belonging to each class is counted and the decision is made whether   the test object belongs to a particular class.

### 2022-03-23 2022/PAVTOneAPI/PAVT_Poster/poster.tex
fix some mistakes

@@
 
-Распараллеливание организовано следующим образом: в ходе выполнения одной итерации алгоритма одновременно проводятся \(P \geq 1\) испытаний в различных точках области поиска.
-Важно отметить, что рассматриваются функции, в которых самым трудоемким и времязатратным этапом работы является это вычисление значения функции в точке.
-Поэтому организация одновременных вычислений этих значений положительно влияет на скорость работы алгоритма для таких функций.
+Распараллеливание организовано следующим образом: в ходе выполнения одной итерации алгоритма одновременно проводятся \(P \geq 1\) испытаний в наиболее перспективных точках области поиска.
+В прикладных задачах глобальной оптимизации проведение даже одного испытания является трудоемкой операцией.
+Поэтому параллельное проведение нескольких испытаний повышает скорость работы алгоритма.
 
@@
         \begin{column}[t]{0.47\paperwidth}
-          \begin{block}{Общая схема алгоритма}
+          \begin{block}{Общая схема параллельного алгоритма}
 
-Общая схема одной итерации одномерного метода:
+Одна итерация состоит в выполнении следующих действий:
               \begin{enumerate}
                 \justifying
-                \item Упорядочить точки предшествующих испытаний в порядке возрастания их координат: \(a=x_{0}<...<x_{i}<...<x_{k}=b\).
-                \item Вычислить для каждого интервала \((x_{i-1};x_{i}),1\leq i\leq k\)  характеристику \(R(i)\) .
-                \item Упорядочить характеристики \(R(i)\) в порядке убывания и выбрать \(P \) наибольших из них с номерами интервалов \(t_{j}, 1\leqslant j \leqslant P: R(t_{1})\geqslant R(t_{2})\geqslant ...
-\geqslant  R(t_{k-1})\geqslant R(t_{k})\)
-                \item Параллельно провести следующие испытания в точках \(x^{k+j}, 1\leqslant j \leqslant P, \) \(x^{k+j}=d(t)\in (x_{t-1};x_{t})\) , где \(d(t)\)  — правило размещения точки следующего испытания в интервале с номером \(t\).
-                \item Проверить выполнение критерия остановки \(x_{t}-x_{t-1}<\varepsilon\).
+                \item Упорядочить точки предшествующих испытаний в порядке возрастания их координат: \(x_{0}<...<x_{i}<...<x_{k}\);
+                \item Для каждого интервала \((x_{i-1}, x_{i}),1\leqslant i\leqslant k,\) вычислить его \textit{характеристику} \(R(i)\);
+                \item Упорядочить характеристики \(R(i)\) в порядке убывания \(R(t_{1})\geqslant R(t_{2})\geqslant ...
+\geqslant  R(t_{k-1})\geqslant R(t_{k})\) и выбрать \(P\) интервалов \((x_{t_j-1}, x_{t_j}), 1\leqslant j \leqslant P,\) с наибольшими характеристиками;
+                \item Провести следующие испытания параллельно в точках  \(x^{k+j} \in (x_{t_j-1}, x_{t_j})\), \(1\leqslant j \leqslant P,\) вычисляемых согласно правилу размещения точки следующего испытания в интервале с заданным номером \(t_j\);
+                \item Проверить выполнение критерия остановки \(\Delta_{t_j}<\varepsilon, 1\leqslant j \leqslant P\).
               \end{enumerate}
@@
 
-Одним из вариантов реализации данного подхода является использование набора инструментов Intel oneAPI.
-Достоинством Intel oneAPI является простота, открытость и возможность легкой интегрирации в существующий код.
+Мы реализовали рассмотренный параллельный алгоритм с использованием набора инструментов Intel oneAPI.
+Достоинством Intel oneAPI является простота, открытость и возможность легкой интеграции в существующий код.
 Intel oneAPI позволяет написать код один раз и в дальнейшем запускать его на различных устройствах.
 
-В рамках данной задачи нам необходимы возможности распараллеливания, которые обеспечиваются за счет включения в oneAPI языка data parallel c++ и набора библиотек, облегчающих межархитектурную разработку.
+При реализации мы использовали возможности распараллеливания, которые обеспечиваются языком Data Parallel C++ и набором библиотек Intel oneAPI, облегчающих межархитектурную разработку.
 
@@
 
-На данный момент получены результаты вычислительных экспериментов, в которых тестовые задачи формируются специальным генератором \cite{GKLS}.
-В таблицах \ref{table:GKLS_RES_1}, \ref{table:GKLS_RES_1} приведены среднее число итераций параллельного алгоритма глобального поиска и ускорение относительно последовательной версии.
+На данный момент получены результаты вычислительных экспериментов на тестовых задачах, формируемых специальным генератором \cite{GKLS}.
+В таблицах \ref{table:GKLS_RES_1}, \ref{table:GKLS_RES_2} приведены среднее число итераций параллельного алгоритма глобального поиска и ускорение относительно последовательной версии.
 В каждом эксперименте решалось 100 задач размерности \(N = 4\) и \(N = 5\).
-Эксперименты проведены на компьютере с процессором Intel Core I5-7300HQ 2.5 GHz с видеокартой HD Graphics 630, 16 GB RAM, использовался язык \textit{data parallel c++}.
+Эксперименты проведены на компьютере с процессором Intel Core I5-7300HQ 2.5 GHz с видеокартой HD Graphics 630, 16 GB RAM, использовался язык \textit{Data Parallel C++}.
 

### 2022-03-24 2022/PAVTOneAPI/PAVT_Poster/poster.tex
fix table caption
Поправлен постер

@@
 \usepackage{fontspec}
+\usepackage{indentfirst}
 
@@
 
+%\bibliographystyle{unsrt}
+
 \begin{document}
@@
     \begin{columns}[t]
-        \begin{column}[t]{0.47\paperwidth}
+        \begin{column}[t]{0.465\paperwidth}
             \begin{block}{Постановка задачи}
+
+            \leftskip=0.5cm \rightskip=0.5cm 
+            \setlength{\parindent}{1.25cm} %
              Задача многомерной многоэкстремальной оптимизации может быть определена как поиск наименьшего значения действительной функции \(\phi(y)\)  в гиперинтервале \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
@@
           \begin{block}{Редукция размерности с помощью кривых Пеано}
-
+\leftskip=0.5cm \rightskip=0.5cm 
+\setlength{\parindent}{1.25cm} %
 В ННГУ им.
@@
 
- \begin{minipage}[t]{.47\textwidth}
+\leftskip=0.02cm 
+\setlength{\parindent}{0.02cm}
+ \begin{minipage}[t]{.98\textwidth}
               \begin{figure}
-                  \centering
-                  \includegraphics[scale=1.49]{images/pean_2.jpg}
+                  %\centering
+                  \center{\includegraphics[width=1.01\textwidth]{images/pean_2 (1)_8.jpg}}
+                  %\includegraphics[scale=1.468]{images/pean_2.jpg}
+                 % \caption{Построение кривой Пеано, заполняющей трехмерный куб}
               \end{figure}
@@
 \begin{block}{Метод глобальной оптимизации}
-
+\leftskip=0.5cm \rightskip=0.5cm 
+\setlength{\parindent}{1.25cm} %
 В процессе работы алгоритма строится последовательность точек \(x^k\), в которых проводятся \textit{испытания}, т.е. вычисляются значения целевой функции \(\phi^k=\phi(y(x^k))\).
@@
 
- \begin{minipage}[t]{.47\textwidth}
+\leftskip=0.001cm 
+\setlength{\parindent}{0.001cm}
+ \begin{minipage}[t]{0.96\textwidth}
               \begin{figure}
-                  \centering
-                  \includegraphics[scale=2.05]{images/PlotGS_1.png}
+                 % \centering
+                 \center{\includegraphics[width=1.045\textwidth]{images/PGS2.eps}}
+                % \includegraphics[scale=1.095]{images/PlotGS_7.png}
+                %  \caption{}
               \end{figure}
@@
           \begin{block}{Общая схема параллельного алгоритма}
-
+\leftskip=0.5cm \rightskip=0.5cm 
+\setlength{\parindent}{1.25cm} %
 Одна итерация состоит в выполнении следующих действий:
               \begin{enumerate}
+              \leftskip=0.5cm \rightskip=0.5cm 
+\setlength{\parindent}{1.25cm} %
                 \justifying
@@
 
+\leftskip=0.5cm \rightskip=0.5cm 
+\setlength{\parindent}{1.25cm} %
 Мы реализовали рассмотренный параллельный алгоритм с использованием набора инструментов Intel oneAPI.
-Достоинством Intel oneAPI является простота, открытость и возможность легкой интеграции в существующий код.
+Достоинствами Intel oneAPI являются простота, открытость и возможность легкой интеграции в существующий код.
 Intel oneAPI позволяет написать код один раз и в дальнейшем запускать его на различных устройствах.
@@
 
+\leftskip=0.5cm \rightskip=0.5cm 
+\setlength{\parindent}{1.25cm} %
 На данный момент получены результаты вычислительных экспериментов на тестовых задачах, формируемых специальным генератором \cite{GKLS}.
@@
     \centering
-    \caption{Число итераций и ускорение при решении серии задач размерности $N=4$}
+    \caption{Результаты решения серии задач размерности $N=4$}
      \renewcommand{\arraystretch}{1.4}
@@
     \centering
-    \caption{Число итераций и ускорение при решении серии задач размерности $N=5$}
+    \caption{Результаты решения серии задач размерности $N=5$}
     \renewcommand{\arraystretch}{1.4}
@@
           \begin{block}{Литература}
+           %\bibliographystyle{unsrt}
+          %\bibliography{bibliography}
+
             \printbibliography

### 2022-04-03 2022/Sensors/sensors.tex
add experiments and conclusion

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Problem Statement}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Dimensionality reduction using space-filling curves}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Constructing a family of Peano curves}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Special features of the implementation of the kNN method using multiple space-filling curves}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Experimental Results}
 
-\begin{specialtable}[H] 
-	\caption{The average number of tests when minimizing Shekel test functions (the number of unsolved problems is indicated in parentheses)}\label{table:average_Shekel}
-	\center
-\begin{tabular}{cccc}
-\toprule
-        & \textbf{$\epsilon = 10^{-4}$} & \textbf{$\epsilon = 10^{-3}$} & \textbf{$\epsilon = 10^{-2}$} \\
-\midrule													
-DIRECT         & 64(1) &  34(6)   & 20(17)    \\
-GSA            & 106  & 53  &  31   \\ 
-GSA-DT         & 49   & 43  &  35   \\
+Let us compare the kNN method that uses multiple evolvents (hereafter referred to as kNN-ME) with the kNN method using KD-tree (hereafter referred to as kNN-KD) \cite{Hou2018} to handle multidimensional data.
 
-\bottomrule
-\end{tabular}
-\end{specialtable}
+The study was performed using several datasets, both randomly generated for testing and real-world  datasets from the database [https://archive.ics.uci.edu/ml/index.php]:
+
+\textbf{DS-Random-1} - a set of random data generated using the random module of the Python NumPy library. The data has a uniform distribution and contains values on an interval from 0 to 1.
+
+\textbf{DS-Random-2} - a set of random data generated by the scikit-learn library (version 1.0.2) function \texttt{sklearn.datasets.make\_classification}.
+
+\textbf{DS-SkinSegmentation} - a dataset of dimension 3 (RGB values) to distinguish human skin colour from other objects in images \cite{ds-1}.
+
+\textbf{DS-CarEvaluation} - a dataset of dimension 6, representing data from a car valuation model based on the values of 6 features \cite{ds-2}.
+
+Two kinds of studies were carried out:
+\begin{itemize}
+\item The time required to find K nearest "neighbours" for a test object was measured;
+\item The accuracy of recognition of test objects was studied, based on the analysis as to whether their neighbours belong to certain classes.
+\end{itemize}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions and Future Work}
+\subsection{Test infrastructure}
+
+The computational experiments were performed on the following test infrastructure (Table \ref{tab:t1}).
+\begin{table}[h]
+	\centering
+    \caption{Test infrastructure}\label{tab:t1}
+	\begin{tabular}{|l|l|}
+		\hline
+		CPU & Intel Core i3-7100 (3.9 HHz) \\
+		\hline
+		RAM & 8 GB \\ 
+		\hline
+		Operating system & Windows 10 \\
+		\hline
+		Compiler & Intel(R) oneAPI DPC++/C++ Compiler, Version 2022.0.0 \\
+		\hline
+	\end{tabular}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Investigation of neighbour search times using test datasets}
+
+Several datasets were synthesized to examine the search time for neighbours. First, DS-Random-1 and DS-Random-2 sets consisting of 100000 and 10000 objects, respectively, were generated.
+
+When generating DS-Random-1, the standard random module from the Python NumPy library was used. The data dimension $N$ varied from 2 to 4; the generated data had a uniform distribution on $[0, 1]$.
+
+To generate the DS-Random-2 set, the \texttt{sklearn.datasets.make\_classification} function from the scikit-learn library (version 1.0.2) was used with the following parameters:
+
+\begin{itemize}
+\item \texttt{n\_samples} = 10000 (number of objects),
+\item \texttt{n\_features} =  $N$ (dimensionality),
+\item \texttt{n\_redundant} = 0 (number of redundant features),
+\item \texttt{n\_informative} = $N - 1$  (number of informative features),
+\item \texttt{n\_clusters\_per\_class} = 1 (number of clusters per class of objects generated).
+\end{itemize}
+
+The running times used by the different methods (kNN-ME, procedure A, kNN-ME, procedure B, kNN-KD) for DS-Random-1 and DS-Random-2 datasets with $N = 2$ are shown in Fig.
+\ref{fig3}.
+The abscissa axis represents the number of nearest neighbours that the algorithm is searching for, the ordinate axis represents the running time of the method.
+
+\begin{figure}
+\widefigure
+\includegraphics[width=1.0\linewidth]{fig3.png}
+\caption{Running times of the algorithms with $N = 2$\label{fig3}}
+\end{figure}
+
+Running times of the methods for nearest neighbours search with $N = 3, 4$ are given in Figs. \ref{fig4}, \ref{fig5} respectively.
+
+\begin{figure}
+\widefigure
+\includegraphics[width=1.0\linewidth]{fig4.png}
+\caption{Running times of the algorithms with $N = 3$\label{fig4}}
+\end{figure}
+
+\begin{figure}
+\widefigure
+\includegraphics[width=1.0\linewidth]{fig5.png}
+\caption{Running times of the algorithms with $N = 4$\label{fig5}}
+\end{figure}
+
+All the methods considered show a small variation in running time when switching from one dimension to another.
+The number of objects being processed slightly increases the search time, and there is also a quasi-linear increase in time as the number of neighbours identified increases.
+All these results are in good agreement with theoretical estimates of the time complexity of the methods under study.
+
+On the whole, the experimental results show that the kNN-ME method based on search without removal (procedure A) works on average 2 times faster than the method that involves removal of elements found   (procedure B).
+Therefore, we will work with the kNN-ME(A) when investigating classification accuracy.
+In fact, both variants of the kNN-ME method are 5-7 times faster (depending on the dataset) than the kNN-KD method.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Investigation of neighbour search times on real-world data}
+
+Let us consider the DS-Skin set taken from the collection \cite{ds-1}.
+The objects in the set belong to two classes: the first class represents points in RGB colour space that belong to human skin images; the second class represents colour coordinates that do not belong to such images.
+
+The number of objects of the first class is about 17000, the number of objects of the second class is 43000. The total number of objects examined is 60000.
+
+\begin{figure}[h]
+\widefigure
+\includegraphics[width=0.5\linewidth]{fig6.png}
+\caption{Comparison of neighbour search times when determining the colour of an image point\label{fig6}}
+\end{figure}
+
+The results of performance comparison of the methods on real-world data are qualitatively similar to those obtained by comparison on test data.
+The kNN-ME method (implementation A) is 1.5 times faster than implementation B, and both implementations are at least 5 times faster than the kNN-KD method.
+
+The performance of kNN-ME and kNN-KD methods was also tested using a real-world dataset \cite{ds-2}.
+There are 4 classes of objects in this dataset of dimension $N = 6$.
+One particular feature of the dataset is the fact that the number of objects belonging to each of the 4 classes differs widely (Table \ref{tab:t2}).
+
+\begin{table}[h]
+	\centering
+    \caption{Classes in the CarEvoluation set}\label{tab:t2}
+	\begin{tabular}{|c|c|}
+		\hline
+		Object class & Number of objects in class \\
+		\hline
+		1 & 1400 \\ 
+		\hline
+		2 & 400 \\
+		\hline
+		3 & 70 \\
+		\hline
+		4 & 65 \\
+		\hline
+	\end{tabular}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Investigation of classification accuracy}
+
+According to the general idea of the kNN method, after finding $K$ nearest neighbours to the test object, an attempt is made to determine the class of the test object.
+The more neighbours belong to a certain class, the higher the probability that the test object belongs to the same class.
+We will choose an odd number of neighbours, so as to avoid a situation when the nearest objects in equal numbers belong to different classes.
+
+Unfortunately, determining the class of the test object is prone to error.
+Errors may occur because of the "location" of the object under test on the hypercube edge in immediate proximity to objects of another class.
+Figure \ref{fig7} shows an example of a random test set where some objects of one class are surrounded by objects of an "alien" class.
+
+\begin{figure}[h]
+\widefigure
+\includegraphics[width=0.5\linewidth]{fig7.png}
+\caption{"Surrounded by aliens"\label{fig7}}
+\end{figure}
+
+DS-Random-2 with $N = 2, 3, 4$ and DS-Skin were used as datasets. The experiment to recognise the class of test objects was carried out as follows.
+
+\begin{itemize}
+\item The source dataset was divided into two samples: a training sample and a test sample. The test sample size was 10\% of the initial number of objects.
+\item $K$ neighbours were searched for each object in the test sample.
+\item The class of an object under test was determined from the set of neighbours found.
+\item The recognition result was checked against the known test object class and an error was recorded if there was a discrepancy with the result. The errors were summed up for the whole test sample.
+\end{itemize}
+
+To divide the source dataset into a training sample and a test sample, we used the \texttt{sklearn.model\_selection.train\_test\_split} function from the scikit-learn library (version 1.0.2).
+
+The results of investigating recognition errors on the synthetically generated random datasets (DS-Random-2) are given in Table \ref{tab:t3}.
+The numerical values in the table show the percentage of recorded recognition errors.
+
+\begin{table}[h]
+    \caption{Percentage of errors in DS-Random-2 recognition}\label{tab:t3}
+	%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
+	\begin{tabular} {|p{0,8cm}|p{1,2cm}|p{0,8cm}|p{0,8cm}|p{1,2cm}|p{0,8cm}|p{0,8cm}|p{1,2cm}|p{0,8cm}|p{0,8cm}|}
+\hline
+K & KDTree (N=2) & A (N=2) & B N=2) & KDTree (N=3) & A (N=3) & B (N=3) & KDTree (N=4) & A (N=4) & B (N=4) \\
+\hline
+5 & 5 & 5 & 5 & 11 & 12 & 12 & 6 & 8 & 8 \\
+\hline
+15 & 4 & 6 & 6 & 10 & 11 & 11 & 10 & 9 & 9 \\
+\hline
+25 & 5 & 7 & 7 & 10 & 11 & 11 & 7 & 9 & 9 \\
+\hline
+35 & 6 & 6 & 6 & 11 & 13 & 13 & 7 & 10 & 10 \\
+\hline
+45 & 7 & 6 & 6 & 10 & 13 & 13 & 7 & 10 & 10 \\
+\hline
+	\end{tabular}
+\end{table}
+
+Based on the results of the experiment with random sets, we can conclude that the kNN-ME method has a recognition accuracy close to that of kNN-KD.
+The error rates of the methods differ, on average, by no more than 2-3\%.
+
+It should be noted that both kNN-ME and kNN-KD demonstrated high error rates, which can be explained by the fact that when generating random data, numerous objects at the class boundaries are generated that are classified incorrectly.
+
+The results of measuring the recognition accuracy on the real-world DS-Skin dataset show that on real-world data both methods show the same percentage of misclassification.
+The specific error percentages obtained with different numbers of neighbours K are shown in Table \ref{tab:t4}.
+
+\begin{table}[h]
+    \centering
+    \caption{Percentage of DS-Skin recognition errors}\label{tab:t4}
+	%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
+	\begin{tabular} {|p{1cm}|p{1,2cm}|p{1cm}|p{1cm}|}
+\hline
+K & KDTree (N=3) & A (N=3) & B (N=3) \\
+\hline
+5 & 0.5 & 0.5 & 0.5 \\
+\hline
+15 & 0.5 & 0.5 & 0.5 \\
+\hline
+25 & 0.5 & 0.5 & 0.5 \\
+\hline
+35 & 1 & 1 & 1 \\
+\hline
+45 & 1 & 1 & 1 \\
+\hline
+	\end{tabular}
+\end{table}
+
+Let us consider the results of determining the class of objects in the CarEvaluation set.
+For objects of classes 1 and 2, we generated test samples containing 20 objects, and for classes 3 and 4, samples containing 5 objects were generated.
+The kNN-ME and kNN-KD methods showed almost identical results: objects from the first two classes were identified with an average error not exceeding 5\%, while objects of classes 3, 4 were identified with an average error of 15\%.
+The average recognition error rates for all classes are shown in Table \ref{tab:t5}.
+
+\begin{table}[h]
+    \centering
+    \caption{Average error rate (in percent) for DS-CarEvoluation recognition}\label{tab:t5}
+	%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
+	\begin{tabular} {|p{1cm}|p{1,2cm}|p{1cm}|p{1cm}|}
+\hline
+K & KDTree (N=6) & A (N=6) & B (N=6) \\
+\hline
+5 & 8 & 11 & 11 \\
+\hline
+7 & 9 & 10 & 10 \\
+\hline
+9 & 8 & 10 & 10 \\
+\hline
+11 & 10 & 11 & 10 \\
+\hline
+13 & 9 & 11 & 11 \\
+\hline
+	\end{tabular}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+
+The paper considers a classical machine learning method, the k nearest neighbour (kNN) method, which is widely used in machine learning practice and has been proven to perform very well because of its "explainability" (the decision to select an object class can be clearly explained).
+The results of practical implementation of one of kNN acceleration approaches based on dimensionality reduction using space-filling curves are presented.
+
+An implementation of the kNN algorithm that uses multiple space-filling curves to better transfer the metric properties of multidimensional space to one-dimensional scales is proposed.
+Theoretical statements are presented about the preservation of the proximity property of objects in multidimensional space and on one of one-dimensional scales when dimensionality reduction is performed.
+
+The software implementation of the kNN method using multiple evolvents (kNN-ME) shows 5-7 times faster performance compared to the algorithm using kd-trees (kNN-KD).
+The comparison was conducted on test and real-world data sets of different size and dimensionality.
+
+Both algorithms demonstrate comparable performance, measured as the percentage of misclassification of objects from the test sample.
+The number of neighbours in the classification was varied from 5 to 100.
+When applied to the synthesized random data both kNN-ME and kNN-KD show a classification error rate of approximately 10\%, while with real-world data both methods show an error rate of about 1\%.
 

### 2022-04-04 2022/Sensors/sensors.tex
Results and Conclusion

@@
 The study was performed using several datasets, both randomly generated for testing and real-world  datasets from the database [https://archive.ics.uci.edu/ml/index.php]:
+\begin{enumerate}
+\item
+\textbf{DS-Random-1} -- a set of random data generated using the random module of the Python NumPy library. The data has a uniform distribution and contains values on an interval from 0 to 1.
 
-\textbf{DS-Random-1} - a set of random data generated using the random module of the Python NumPy library. The data has a uniform distribution and contains values on an interval from 0 to 1.
+\item
+\textbf{DS-Random-2} -- a set of random data generated by the scikit-learn library (version 1.0.2) function \texttt{sklearn.datasets.make\_classification}.
 
-\textbf{DS-Random-2} - a set of random data generated by the scikit-learn library (version 1.0.2) function \texttt{sklearn.datasets.make\_classification}.
+\item
+\textbf{DS-SkinSegmentation} -- a dataset of dimension 3 (RGB values) to distinguish human skin colour from other objects in images \cite{ds-1}.
 
-\textbf{DS-SkinSegmentation} - a dataset of dimension 3 (RGB values) to distinguish human skin colour from other objects in images \cite{ds-1}.
-
-\textbf{DS-CarEvaluation} - a dataset of dimension 6, representing data from a car valuation model based on the values of 6 features \cite{ds-2}.
+\item
+\textbf{DS-CarEvaluation} -- a dataset of dimension 6, representing data from a car valuation model based on the values of 6 features \cite{ds-2}.
+\end{enumerate}
 
@@
 \begin{itemize}
-\item The time required to find K nearest "neighbours" for a test object was measured;
+\item The time required to find K nearest neighbours for a test object was measured;
 \item The accuracy of recognition of test objects was studied, based on the analysis as to whether their neighbours belong to certain classes.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Test infrastructure}
+%\subsection{Test infrastructure}
 
 The computational experiments were performed on the following test infrastructure (Table \ref{tab:t1}).
-\begin{table}[h]
-	\centering
+
+\begin{specialtable}[H] 
     \caption{Test infrastructure}\label{tab:t1}
-	\begin{tabular}{|l|l|}
-		\hline
+	\center
+\begin{tabular}{ll}
+\toprule
+	  Computer specifications & \\
+\midrule													
 		CPU & Intel Core i3-7100 (3.9 HHz) \\
-		\hline
 		RAM & 8 GB \\ 
-		\hline
 		Operating system & Windows 10 \\
-		\hline
 		Compiler & Intel(R) oneAPI DPC++/C++ Compiler, Version 2022.0.0 \\
-		\hline
+\bottomrule
 \end{tabular}
-\end{table}
+\end{specialtable}
 
@@
 On the whole, the experimental results show that the kNN-ME method based on search without removal (procedure A) works on average 2 times faster than the method that involves removal of elements found (procedure B).
-Therefore, we will work with the kNN-ME(A) when investigating classification accuracy.
 In fact, both variants of the kNN-ME method are 5-7 times faster (depending on the dataset) than the kNN-KD method.
@@
 The kNN-ME method (implementation A) is 1.5 times faster than implementation B, and both implementations are at least 5 times faster than the kNN-KD method.
+Therefore, we will work with the kNN-ME(A) when investigating classification accuracy.
 
@@
 
-\begin{table}[h]
-	\centering
-    \caption{Classes in the CarEvoluation set}\label{tab:t2}
-	\begin{tabular}{|c|c|}
-		\hline
+\begin{specialtable}[H] 
+  \caption{Classes in the CarEvaluation set}\label{tab:t2}
+	\center
+\begin{tabular}{cc}
+\toprule
 	  Object class & Number of objects in class \\
-		\hline
+\midrule													
 		1 & 1400 \\ 
-		\hline
 		2 & 400 \\
-		\hline
 		3 & 70 \\
-		\hline
 		4 & 65 \\
-		\hline
+\bottomrule
 \end{tabular}
-\end{table}
+\end{specialtable}
 
@@
 Unfortunately, determining the class of the test object is prone to error.
-Errors may occur because of the "location" of the object under test on the hypercube edge in immediate proximity to objects of another class.
-Figure \ref{fig7} shows an example of a random test set where some objects of one class are surrounded by objects of an "alien" class.
+Errors may occur because of the ``location'' of the object under test on the hypercube edge in immediate proximity to objects of another class.
+Figure \ref{fig7} shows an example of a random test set where some objects of one class are surrounded by objects of an ``alien'' class.
 
@@
 \widefigure
-\includegraphics[width=0.5\linewidth]{fig7.png}
-\caption{"Surrounded by aliens"\label{fig7}}
+\includegraphics[width=0.7\linewidth]{fig7.png}
+\caption{``Surrounded by aliens''\label{fig7}}
 \end{figure}
@@
 
-\begin{itemize}
+\begin{enumerate}
 \item The source dataset was divided into two samples: a training sample and a test sample. The test sample size was 10\% of the initial number of objects.
@@
 \item The recognition result was checked against the known test object class and an error was recorded if there was a discrepancy with the result. The errors were summed up for the whole test sample.
-\end{itemize}
+\end{enumerate}
 
@@
 
-\begin{table}[h]
+\begin{specialtable}[H] 
   \caption{Percentage of errors in DS-Random-2 recognition}\label{tab:t3}
-	%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-	\begin{tabular} {|p{0,8cm}|p{1,2cm}|p{0,8cm}|p{0,8cm}|p{1,2cm}|p{0,8cm}|p{0,8cm}|p{1,2cm}|p{0,8cm}|p{0,8cm}|}
-\hline
-K & KDTree (N=2) & A (N=2) & B N=2) & KDTree (N=3) & A (N=3) & B (N=3) & KDTree (N=4) & A (N=4) & B (N=4) \\
-\hline
-5 & 5 & 5 & 5 & 11 & 12 & 12 & 6 & 8 & 8 \\
-\hline
-15 & 4 & 6 & 6 & 10 & 11 & 11 & 10 & 9 & 9 \\
-\hline
-25 & 5 & 7 & 7 & 10 & 11 & 11 & 7 & 9 & 9 \\
-\hline
-35 & 6 & 6 & 6 & 11 & 13 & 13 & 7 & 10 & 10 \\
-\hline
-45 & 7 & 6 & 6 & 10 & 13 & 13 & 7 & 10 & 10 \\
-\hline
+	\center
+	\begin{tabular}{ccccccccc}
+\toprule
+K & \multicolumn{2}{c}{ $N=2$ } & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} \\
+		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \noalign{\smallskip}
+  & kNN-KD & kNN-ME & & kNN-KD & kNN-ME & & kNN-KD & kNN-ME \\
+\midrule													
+5  & 5 & 5 & & 11 & 12 & & 6 & 8 \\
+15 & 4 & 6 & & 10 & 11 & & 10 & 9 \\
+25 & 5 & 7 & & 10 & 11 & & 7 & 9 \\
+35 & 6 & 6 & & 11 & 13 & & 7 & 10 \\
+45 & 7 & 6 & & 10 & 13 & & 7 & 10 \\
+\bottomrule
 \end{tabular}
-\end{table}
+\end{specialtable}
 
@@
 
-\begin{table}[h]
-    \centering
-    \caption{Percentage of DS-Skin recognition errors}\label{tab:t4}
-	%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-	\begin{tabular} {|p{1cm}|p{1,2cm}|p{1cm}|p{1cm}|}
-\hline
-K & KDTree (N=3) & A (N=3) & B (N=3) \\
-\hline
-5 & 0.5 & 0.5 & 0.5 \\
-\hline
-15 & 0.5 & 0.5 & 0.5 \\
-\hline
-25 & 0.5 & 0.5 & 0.5 \\
-\hline
-35 & 1 & 1 & 1 \\
-\hline
-45 & 1 & 1 & 1 \\
-\hline
+\begin{specialtable}[H] 
+  \caption{Percentage of DS-Skin recognition errors, $N=3$}\label{tab:t4}
+	\center
+\begin{tabular}{ccc}
+\toprule
+K & kNN-KD & kNN-ME \\
+\midrule													
+5 & 0.5 & 0.5 \\
+15 & 0.5 & 0.5 \\
+25 & 0.5 & 0.5\\
+35 & 1 & 1\\
+45 & 1 & 1\\
+\bottomrule
 \end{tabular}
-\end{table}
+\end{specialtable}
 
@@
 
-\begin{table}[h]
-    \centering
-    \caption{Average error rate (in percent) for DS-CarEvoluation recognition}\label{tab:t5}
-	%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-	\begin{tabular} {|p{1cm}|p{1,2cm}|p{1cm}|p{1cm}|}
-\hline
-K & KDTree (N=6) & A (N=6) & B (N=6) \\
-\hline
-5 & 8 & 11 & 11 \\
-\hline
-7 & 9 & 10 & 10 \\
-\hline
-9 & 8 & 10 & 10 \\
-\hline
-11 & 10 & 11 & 10 \\
-\hline
-13 & 9 & 11 & 11 \\
-\hline
+\begin{specialtable}[H] 
+  \caption{Average error rate (in percent) for DS-CarEvaluation recognition, $N = 6$}\label{tab:t5}
+	\center
+\begin{tabular}{ccc}
+\toprule
+K & kNN-KD & kNN-ME \\
+\midrule													
+5 & 8 & 11 \\
+7 & 9 & 10 \\
+9 & 8 & 10 \\
+11 & 10 & 11 \\
+13 & 9 & 11 \\
+\bottomrule
 \end{tabular}
-\end{table}
+\end{specialtable}
 
@@
 
-The paper considers a classical machine learning method, the k nearest neighbour (kNN) method, which is widely used in machine learning practice and has been proven to perform very well because of its "explainability" (the decision to select an object class can be clearly explained).
+The paper considers a classical machine learning method, the k nearest neighbour (kNN) method, which is widely used in machine learning practice and has been proven to perform very well because of its ``explainability'' (the decision to select an object class can be clearly explained).
 The results of practical implementation of one of kNN acceleration approaches based on dimensionality reduction using space-filling curves are presented.

### 2022-04-06 2022/Sensors/sensors.tex
add one more figure and comment

@@
 
+We employed in our experiments the kNN-KD method from the scikit-learn library; this method was used with default settings.
+For our kNN-ME method, we used the evolvent density $m=12$, the number of evolvents simultaneously used was $L=10$ for datasets 1 -- 3 and $L=7$ for the last dataset.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{specialtable}
+
+The time required to find K nearest neighbours on the CarEvoluation dataset is shown in Fig.~\ref{fig6a}. In this experiment, the number of evolvents used simultaneously was 7. 
+\begin{figure}
+\widefigure
+\includegraphics[width=0.5\linewidth]{fig6a.png}
+\caption{Comparison of neighbour search times for the CarEvoluation dataset\label{fig6a}}
+\end{figure}
+The results of the comparison on this dataset also demonstrate that the kNN-ME method is faster compared to the kNN-KD method.
 

### 2022-04-06 2022/Sensors/sensors.tex
fix typo in conclusion

@@
 
-The paper considers a classical machine learning method, the k nearest neighbour (kNN) method, which is widely used in machine learning practice and has been proven to perform very well because of its ``explainability'' (the decision to select an object class can be clearly explained).
+The paper considers a classical machine learning method, the k nearest neighbour (kNN) method, which is widely used in machine learning practice and has been proven to perform very well because of its ``explainability'' (the decision  regarding the choice of an object class can be clearly explained).
 The results of practical implementation of one of kNN acceleration approaches based on dimensionality reduction using space-filling curves are presented.

### 2022-04-09 2022/RSD/paper.tex
LaTeX template for the paper

--- 2018-12-28 2019/MOTOR/motor.tex
@@
 % LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
+% Version 2.21 of 2022/01/12
 %
 \documentclass[runningheads]{llncs}
+%
+\usepackage[T1]{fontenc}
+% T1 fonts will be used to generate the final print and online PDFs,
+% so please use T1 fonts in your manuscript whenever possible.
+% Other font encondings may result in incorrect characters.
 %
@@
 %
-% If you use the hyperref package, please uncomment the following line
+% If you use the hyperref package, please uncomment the following two lines
 % to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
+%\usepackage{color}
+%\renewcommand\UrlFont{\color{blue}\rmfamily}
+%
 \begin{document}
 %
-\title{ Generalized nested optimization scheme for global optimization problems
-\thanks{This study was supported by the Russian Science Foundation, project 
-No.\,16-11-10150.}}
+\title{Contribution Title\thanks{This study was supported by the Russian Science Foundation, project No.\,21-11-00204}}
 %
@@
 %
-\author{Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
-Ilia Lebedev\orcidID{0000-0002-8736-0652}}
+\author{
+Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
+Ilya Lebedev \orcidID{0000-0002-8736-0652} \and
+Denis Karchkov \orcidID{}
+}
 %
-\authorrunning{K. Barkalov, I. Lebedev}
+\authorrunning{K. Barkalov, I. Lebedev, D. Karchkov}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
-\email{konstantin.barkalov@itmm.unn.ru}}
+\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
+\email{konstantin.barkalov@itmm.unn.ru}, 
+\email{ilya.lebedev@itmm.unn.ru}}
+
 %
@@
 \section{First Section}
-\subsection{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is not indented. The first paragraph that follows a table, figure,
-equation etc. does not need an indent, either.
 
-Subsequent paragraphs, however, are indented.
+\cite{Strongin2000}
 
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
+\subsubsection{Acknowledgements} Please place your acknowledgments at
+the end of the paper, preceded by an unnumbered run-in heading (i.e.
+3rd-level heading).
 
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
-
-%\begin{figure}
-%\includegraphics[width=\textwidth]{fig1.eps}
-%\caption{A figure caption is always placed below the illustration.
-%Please note that short captions are centered, while long ones are
-%justified by the macro package automatically.} \label{fig1}
-%\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{globalizerSystem}, a
-book~\cite{strSergGO}, proceedings without editors~\cite{Jones2009}.
 %

### 2022-04-11 2022/Sensors/sensors.tex
add new fig files with .pdf extension
fix some typos

--- 2022-04-06 2022/Sensors/sensors.tex
@@
 \abstract{
-The paper considers a time-efficient implementation of the k-nearest neighborhood (kNN) algorithm.
+The paper considers a time-efficient implementation of the $k$ nearest neighbours (kNN) algorithm.
 A well-known approach for accelerating the kNN algorithm is to utilize dimensionality reduction methods based on the use of space-filling curves.
@@
 assigning the object under study to one of the available classes.
-One of the well-known classification methods is the kNN (k-nearest neighbourhood) method.
+One of the well-known classification methods is the kNN ($k$ nearest neighbours) method.
 Although the earliest papers related to this method appeared more than $50$ years ago \cite{Cover1967, Cover1968}, theoretical studies of various aspects of using kNN are still going on \cite{Pestov2013,Mirkes2020}.
@@
 The paper is organized as follows.
-Section 2 describes the scheme for constructing a Peano-Hilbert curve approximation (evolvent) and the kNN method, which uses the evolvent to reduce the dimensionality of the data.
+Section \ref{ps} contains the problem statement and description of the kNN method.
+Section \ref{Peano} describes the scheme for constructing a Peano-Hilbert curve approximation (evolvent) and the kNN method, which uses the evolvent to reduce the dimensionality of the data.
 The fundamental drawbacks of this dimensionality reduction method are noted.
-Section 3 proposes a way to overcome the drawbacks mentioned, based on the use of multiple evolvents.
+Section \ref{MPeano} proposes a way to overcome the drawbacks mentioned, based on the use of multiple evolvents.
 A theoretical statement is formulated about preserving some proximity information in the multidimensional space on one of the one-dimensional scales.
-A kNN scheme using multiple evolvents is presented.
-Section 4 contains experimental results comparing different implementations of kNN using synthetic and real-life data.
-Section 5 concludes the paper.
+A kNN scheme using multiple evolvents is presented in Section \ref{knnme}.
+Section \ref{results} contains experimental results comparing different implementations of kNN using synthetic and real-life data.
+Section \ref{conclusions} concludes the paper.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Problem Statement}
+\section{Problem Statement}\label{ps}
 
@@
 where the value of the $j$-th vector coordinate corresponds to the numerical value of the $j$-th feature of the object $\omega$.
-
 We will also assume that there is already some number of objects exactly classified (a test sample), i.e. for each object we know which class it belongs to.
@@
 The $k$ nearest neighbours (kNN) method is based on the following simple rule: an object is considered to belong to the class to which most of its nearest neighbours belong.
-Here, "neighbours" refers to objects that are close (in one sense or another) to the one being studied.
+Here, ``neighbours'' refers to objects that are close (in one sense or another) to the one being studied.
 Note that in order to apply this method, some metric $L(\omega_i, \omega_j)$, i.e.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Dimensionality reduction using space-filling curves}
+\section{Dimensionality reduction using space-filling curves}\label{Peano}
 
@@
 This can be done by excluding $x_i$ from the set being searched or by comparing the found $x_i$ with the previously found ones to find a match.
-For example, one possible implementation of searching for $K$ nearest neighbours could be as follows: in relation to $x^*$, we look for the nearest values to the left and right on the one-dimensional scale, choose the minimum one and continue searching sideways until exactly K neighbours are found.
+For example, one possible implementation of searching for $K$ nearest neighbours could be as follows: in relation to $x^*$, we look for the nearest values to the left and right on the one-dimensional scale, choose the minimum one and continue searching sideways until exactly $K$ neighbours are found.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Constructing a family of Peano curves}
+\section{Constructing a family of Peano curves}\label{MPeano}
 
@@
 
+\section{Implementation features of the kNN method using multiple space-filling curves}\label{knnme}
+
 The scheme of implementation of the kNN algorithm using multiple Peano curves will look as follows.
-
 For every object $\omega_i$ a set of its preimages on different one-dimensional scales is assigned 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Special features of the implementation of the kNN method using multiple space-filling curves}
+%\subsection{Implementation features of the kNN method using multiple space-filling curves}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experimental Results}
+\section{Experimental Results}\label{results}
 
@@
 
-The study was performed using several datasets, both randomly generated for testing and real-world  datasets from the database [https://archive.ics.uci.edu/ml/index.php]:
+The study was performed using several datasets, both randomly generated for testing and real-world  datasets from the database \url{https://archive.ics.uci.edu/ml/index.php}:
 \begin{enumerate}
@@
 \begin{itemize}
-\item The time required to find K nearest neighbours for a test object was measured;
+\item The time required to find $K$ nearest neighbours for a test object was measured;
 \item The accuracy of recognition of test objects was studied, based on the analysis as to whether their neighbours belong to certain classes.
@@
 \widefigure
-\includegraphics[width=1.0\linewidth]{fig3.png}
+\includegraphics[width=1.0\linewidth]{fig3.pdf}
 \caption{Running times of the algorithms with $N = 2$\label{fig3}}
@@
 \widefigure
-\includegraphics[width=1.0\linewidth]{fig4.png}
+\includegraphics[width=1.0\linewidth]{fig4.pdf}
 \caption{Running times of the algorithms with $N = 3$\label{fig4}}
@@
 \widefigure
-\includegraphics[width=1.0\linewidth]{fig5.png}
+\includegraphics[width=1.0\linewidth]{fig5.pdf}
 \caption{Running times of the algorithms with $N = 4$\label{fig5}}
@@
 \widefigure
-\includegraphics[width=0.5\linewidth]{fig6.png}
+\includegraphics[width=0.5\linewidth]{fig6.pdf}
 \caption{Comparison of neighbour search times when determining the colour of an image point\label{fig6}}
@@
 
-The time required to find K nearest neighbours on the CarEvoluation dataset is shown in Fig.~\ref{fig6a}. In this experiment, the number of evolvents used simultaneously was 7. 
+The time required to find $K$ nearest neighbours on the CarEvoluation dataset is shown in Fig.~\ref{fig6a}. In this experiment, the number of evolvents used simultaneously was 7. 
 \begin{figure}
 \widefigure
-\includegraphics[width=0.5\linewidth]{fig6a.png}
+\includegraphics[width=0.5\linewidth]{fig6a.pdf}
 \caption{Comparison of neighbour search times for the CarEvoluation dataset\label{fig6a}}
@@
 \toprule
-K & \multicolumn{2}{c}{ $N=2$ } & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} \\
+$K$ & \multicolumn{2}{c}{ $N=2$ } & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} \\
 		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \noalign{\smallskip}
@@
 The results of measuring the recognition accuracy on the real-world DS-Skin dataset show that on real-world data both methods show the same percentage of misclassification.
-The specific error percentages obtained with different numbers of neighbours K are shown in Table \ref{tab:t4}.
+The specific error percentages obtained with different numbers of neighbours $K$ are shown in Table \ref{tab:t4}.
 
@@
 \toprule
-K & kNN-KD & kNN-ME \\
+$K$ & kNN-KD & kNN-ME \\
 \midrule													
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
+\section{Conclusions}\label{conclusions}
 
-The paper considers a classical machine learning method, the k nearest neighbour (kNN) method, which is widely used in machine learning practice and has been proven to perform very well because of its ``explainability'' (the decision to select an object class can be clearly explained).
+The paper considers a classical machine learning method, the $k$ nearest neighbours (kNN) method, which is widely used in machine learning practice and has been proven to perform very well because of its ``explainability'' (the decision  regarding the choice of an object class can be clearly explained).
 The results of practical implementation of one of kNN acceleration approaches based on dimensionality reduction using space-filling curves are presented.
@@
 
-\dataavailability{The data analyzed in this study are openly available at https:// .}
+\dataavailability{The data analyzed in this study are openly available at\\ \url{https://archive.ics.uci.edu/ml/index.php}.}
 

### 2022-04-12 2022/RSD/paper.tex
fix a lot of LaTeX errors

@@
 \documentclass[runningheads]{llncs}
+
+\usepackage{hyperref}
 %
@@
 % to display URLs in blue roman font according to Springer's eBook style:
-%\usepackage{color}
-%\renewcommand\UrlFont{\color{blue}\rmfamily}
+\usepackage{color}
+\renewcommand\UrlFont{\color{blue}\rmfamily}
+\usepackage{amsmath} %?????????!!!!!!!
 %
@@
 %
-\title{Contribution Title\thanks{This study was supported by the Russian Science Foundation, project No.\,21-11-00204}}
+\title{Analysis and elimination of bottlenecks in parallel algorithm for solving global optimization problems}
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Analysis and elimination of bottlenecks in parallel algorithm}
 % If the paper title is too long for the running head, you can set
@@
 Ilya Lebedev \orcidID{0000-0002-8736-0652} \and
-Denis Karchkov \orcidID{}
+Denis Karchkov \orcidID{0000-0002-4164-5927}
 }
@@
 \institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
-\email{konstantin.barkalov@itmm.unn.ru}, 
-\email{ilya.lebedev@itmm.unn.ru}}
+\email{\{konstantin.barkalov,ilya.lebedev,karchkov\}@itmm.unn.ru}
+}
 
@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-150--250 words.
+The paper presents the results of investigation of efficiency of the parallelization of a global optimum search algorithm.
+Two stages of solving the problem were distinguished: the search of a rough approximation to the global solution and its local refining.
+The novelty of the parallel algorithm considered in the present pa-per consists in the parallelization of the whole computational process – both at the global search stage and at the one of the local refining of the solution found.
+The theoretical estimates of the algorithm parallelization efficiency are presented.
+The computational experiments were carried out.
+The results of the ones confirmed the theoretical conclusions.
 
@@
 %
-\section{First Section}
+\section{Introduction}
 
-\cite{Strongin2000}
+The paper is devoted to the investigation of the parallel algorithms for solving complex optimization problems arising in various fields of science and technology.
+Several examples of such problems are presented in the review \cite{Pinter2006}.
 
-\subsubsection{Acknowledgements} Please place your acknowledgments at
-the end of the paper, preceded by an unnumbered run-in heading (i.e.
-3rd-level heading).
+As a rule, the objective function $\varphi(y)$ in such problems is a ``black-box''-one, i.e.
+it is defined not by a formula but as a software-implemented algorithm for computing its values at the points of the search domain $D$.
+In this case, the analytical methods for the search of the exact problem solution is $y^{*} \in D$, and its numerical solving is reduced to the construction of an approximate solution $y_k^* \in D, \|y^{*} - y_k^*\| \leq \varepsilon$, where $\varepsilon > 0$ is a predefined precision.
+Hereafter, the process of computing the objective function at some point will be called a search trial.
+
+The problem of search for the values of the parameters of complex mathematical models from the experimental data is a typical example of a global optimization problem with a ``black-box'' objective function.
+Among such problems, there are, for example, the inverse problems of chemical kinetics \cite{Akhmadullina2017,Nurislamova2016}.
+The number of parameters in such models can be several tens that requires using the efficient parallel global optimization algorithms for solving these ones.
+
+From the viewpoint of mathematics, the inverse problem is a global optimization one, i.e.
+the problem of search for the global minimizer of a function on some set.
+Formally, the point $y^{*} \in D$ is called the global minimizer for the function $\varphi(y)$ on the set $D$ if for all $y \in D$ the inequality $\varphi(y^{*}) \leq \varphi(y)$ holds.
+
+The solution of a global optimization problem always implies performing the search trials at the nodes of some grid (uniform, random, or adaptive one) in the search domain.
+The number of nodes in the grid increases exponentially with increasing dimensionality of the problem being solved.
+At the same time, to find the local extremum, the function values are computed in the nodes constituting not a grid but some trajectory starting from the initial guess point and ending at the solution point.
+
+Formally, the point $y^{\prime} \in D$ is called the local minimizer for a function on the set $D$ if there exists such a number $\varepsilon > 0$ that for all $y \in D$ such that $\|y - y^{\prime}\| < \varepsilon$, the inequality $\varphi(y^{\prime}) \leq \varphi(y)$ holds.
+Since one needs to explore not the whole search domain $D$ to find a local minimizer, and, on the other hand, the descend trajectory is, generally speaking, a one-dimension object, the computational costs for the local optimization methods are smaller essentially than the ones for the global search.
+One can say that the local minimum search problem consists in finding the local minimizer $y^{*}(y^{0})$ for given initial guess point $y^{0} \in D$, which falls into the attraction region of $y^{*}$.
+
+There are many cases requiring the search for the local optimal solutions.
+One of such cases arises when an estimate of the globally-optimal solution $y_k^*$ is obtained with a non-satisfactory precision at some $k^{th}$ iteration of the global optimization method.
+In this case, one can find a locally–optimal solution $y_{loc}^*$, (corresponding to the initial guess point $y_k^*$) with a high precision.
+The local minimum $y_{loc}^*$ found depends, generally speaking, on $y_k^*$, i.e.
+$y_{loc}^*(y_k^*)$ will be more precise solution of the global optimization problem.
+
+As a rule, the computational costs for the local refining of the solution is smaller essentially than the costs, which would be necessary for the multiextremal optimization procedures to achieve the same precision as the local method can ensure.
+How-ever, in the case of using the parallel global optimization algorithm (at the first stage) and sequential local method (at the second stage), the work times of these ones can become comparable.
+However, the parallelization efficiency would be low.
+
+The novelty of the parallel algorithm considered in the present paper consists in the parallelization of the whole computing process – both at the global search stage and at the one of the local refining of the solution found.
+This approach can be relevant for the problems, which the objective functions are gully-wise with a weak multiextremality.
+In such problems, the stage of searching the rough approximation of the global solution and the one of its local refining are comparable in the computation costs.
+For example, in the work \cite{Gubaydullin2021} the global search stage only was parallelized; the local refining was made with a sequential algorithm.
+
+The main part of the paper has the following structure.
+In Section 2, the global optimization problem statement is presented and the properties of the objective function allowing developing the efficient global minimum search algorithms are described.
+In Section 3, the scheme of the parallel global optimization algorithm is presented and the theoretical estimates of its speedup are given.
+In Section 4, two local optimization methods (BFGS method and Hooke-Jeeves one) are considered, which can be applied for refining the solution found.
+Possible parallelization schemes for these methods are discussed.
+Section 5 contains the results of the computational experiments conducted using both the test problems and the applied ones.
+Section 6 concludes the paper.
+
+\section{Global optimization problem statement}
+
+Let us consider a problem of searching the global minimum of an $N$-dimensional function $\varphi(y)$ in a hyperinterval $D$
+\begin{equation}\label{problemN}
+\varphi(y^*) = \min{\{ \varphi(y):y \in D \}}
+\end{equation}
+\begin{equation}\label{D}
+D = \{ y \in R^{N}: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}.
+\end{equation}
+
+The main assumption, which the algorithm considered are based on is the fulfillment of Lipschitz condition for the function $\varphi(y)$, i.e.
+$$ \mid \varphi(y_1) - \varphi(y_2) \mid \leq L\|y_1 - y_2\|, y_1, y_2 \in D, 0 < L < \infty. \eqno(3)$$
+Note that the Lipschitz constant $L$ featuring the rate of variation of the function is not known \textit{a priori}.
+This assumption is true for a wide class of problems and is typical of many optimization techniques (see, e.g., \cite{Evtushenko2013,Jones2009,Paulavicius2014}).
+%of finding the optimal values of the parameters of mathematical models from the experimental data.
+
+The known approaches to solving the Lipschitz global optimization problems are based on a generalization of the efficient algorithms for solving the one-dimensional problems onto solving the multidimensional ones (see, for example, the methods of diagonal partitions \cite{Sergeyev2017} or simplicial partitions \cite{PaulaviciusZilinskas2014}).
+In the present work, we have also used the approach reducing the solving of a multidimensional problem to solving corresponding one-dimensional one.
+This approach is based on the idea of the dimensionality reduction using Peano-Hilbert curve $y(x)$ mapping the interval $[0,1]$ of the real axis onto an $N$-dimensional cube
+$$ \{ y \in R^{N}: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N\} = \{y(x): 0 \leq x \leq 1 \}.    \eqno(4)$$
+
+In the theory, the Peano-Hilbert curve is a limiting object.
+Therefore, one needs to construct an approximation to the theoretical curve for the practical application.
+The issues of numerical construction of such approximation (also called \textit{evolvents}) for given dimensionality with given precision were considered in \cite{Sergeyev2013}.
+
+The use of the space-filling curves allows reducing the multidimensional problem \ref{problemN} to a one-dimensional problem 
+$$\varphi(y^*) = \varphi(y(x^*)) = \min{\{ \varphi(y(x)): x \in [0, 1] \}}. \eqno(5)$$
+and applying the efficient one-dimensional optimization methods to solve this one.
+Keeping the property of limited relative differences of the objective function is an important property of using the mappings of this kind.
+If the function $\varphi(y)$ in the domain $D$ satisfies Lipschitz condition, the function $\varphi(y(x))$ in the interval $[0,1]$ will satisfy uniform H{\"o}lder condition
+$$\mid \varphi(y(x_1)) - \varphi(y(x_2)) \mid \leq H {\mid x_1 - x_2 \mid}^{1/N}, x_1, x_2 \in [0,1], \eqno(6)$$
+where H{\"o}lder constant $H$ is related to Lipschitz constant $L$ by the relation $H = 2L\sqrt{N+3}$.
+Therefore, without liming the generality, one can consider the minimization of the one-dimensional function
+$$ f(x) = \varphi(y(x)), x \in [0,1], \eqno(7)$$
+satisfying H{\"o}lder condition.
+
+The algorithm of solving this problem being considered implies the construction of a series of points $x_k$, which the values of the objective function  $z_k = f(x_k)$ are computed in.
+We will divide the solving of the problem into two stages.
+At the first stage, the search of a rough approximation to the global minimum point will be per-formed using the global search algorithm.
+At the second stage, a local refining of the solution found will be conducted with a high precision.
+The methods applied at the first stage and at the second one are described in Section 3 and Section 4, respectively.
+
+\section{Parallel global search algorithm}
+\subsection{Computational scheme of the algorithm}
+
+We will consider the class of problems, in which the time of executing a single search trial doesn’t depend on the point of the trial execution.
+In this case, one can use the synchronous parallelization scheme, which implies $P$ trials to be performed in parallel and synchronously within single method iteration.
+In this case, total number of trials executed after $n$ parallel iterations will be $k=nP$.
+
+The first iteration of the method is performed in a special way.
+At this one, the trials are performed at the boundary points $x^0 = 0$ and $x^1 = 1$ as well as at $(P-2)$ arbitrary different internal points $x^2, x^3, ..., x^P$ in the interval $[0,1]$.
+
+At the second iteration and at all the next ones, the following operations are performed.
+
+Step 1. Renumber the points of the previous trials $\{x^0,...,x^k\}$ by the lower indices in the order of increasing the coordinate values, i.e. 
+$$ 0 = x_0 < x_1 < ... < x_k = 1.$$
+Step 2. Calculate the magnitudes 
+$$ \mu = \max\limits_{1 \leq i \leq k} \frac{\mid z_i - z_{i-1} \mid}{\Delta_i}, $$
+
+\begin{displaymath}
+M = \left\{ \begin{array}{ll}
+                r\mu, & \textrm{$\mu > 0$,}\\
+                1, & \textrm{$\mu = 0$}.
+  \end{array} \right.
+\end{displaymath}
+where $z_i = f(y(x_i)),$  $\Delta_i = {(x_i - x_{i-1})}^{1/N}$, and $r > 1$ is a predefined parameter of the method.
+
+Step 3. For each interval $(x_{i-1}, x_i)$, $1 \leq i \leq k$ compute the characteristic according to the formula
+$$ R(i) = \Delta_i + \frac{(z_i - z_{i-1})^2}{M^2 \Delta_i} - 2 \frac{z_i + z_{i-1}}{M}, 1 \leq i \leq k. $$
+
+Step 4. Arrange the characteristics $R(i)$, $1 \leq i \leq k$ in the decreasing order
+$$ R(t_1) \geq R(t_2) \geq ... \geq R(t_{k-1}) \geq R(t_{k+1}) $$
+and select $P$ largest characteristics with the interval indices $t_j$, $1 \leq j \leq P$.
+
+Step 5. Perform new trials at the points $x^{k+j}$, $1 \leq j \leq P$ computed by the formulae
+$$ x^{k+1} = \frac{x_{t_j} + x_{t_j - 1}}{2} - sign(z_{t_j} - z_{t_j - 1}) \frac{1}{2r} {\Bigg[\frac{\mid z_{t_j} - z_{t_j - 1} \mid}{\mu}\Bigg]}^{N}, 1 < t_j < k+1.$$
+Put the results of the trials performed into the information database of the algo-rithm.\\
+
+Step 6. Check the stopping criterion $\Delta_{t_j} \leq \varepsilon$ where $t_j$ from \cite{} and $\varepsilon > 0$ is the predefined precision.
+
+After completing the algorithm work, take as the estimate of the solution of the problem \ref{problemN} 
+$$ f_k^* = \min\limits_{1 \leq i \leq k} f(x^i), x_k^* = \arg \min\limits_{1 \leq i \leq k} f(x^i).$$
+
+The theoretical substantiation of the algorithm presented here see in more details in \cite{}.
+Here, note briefly that the characteristics of the intervals used in the algorithm \cite{} can be considered as a quantitative measure how particular interval is promising from the viewpoint of finding the global minimum inside this one.
+The inequalities \cite{} arrange the intervals according to the characteristics on the ones, and the trials are performed in parallel in p most promising intervals.
+
+\subsubsection{Theoretical estimates of the speedup of the parallel algorithm}
+Let us describe (according to \cite{}) the theoretical properties of the parallel algorithm featuring its speedup.
+In the problems being solved, the time of executing a single search trial exceeds the time of processing its results essentially.
+Therefore, the speedup in the number of trials is the key characteristic of the efficiency of the parallel global search algorithm
+$$s(p) = k(1)\frac{p}{k(p)}, \eqno(26)$$
+where $k(1)$ is the number of trials executed by the sequential method and $k(p)$ is the number of trials executed by the parallel method employing p parallel processes.\\
+
+Obviously, the quantities of trials $n(p)$ for the algorithms with different degrees of parallelization $p$ will differ from each other.
+Indeed, the sequential algorithm has the full information obtained at the previous $k$ iterations when selecting the point $x^{k+1}$ of the next $(k+1)^{th}$ trial.
+The parallel algorithm selects not one but $p$ points $x^{k+j}$, $1 \leq j \leq p$ on the base of the same information at the $(k+1)^{th}$ iteration.
+It means that the selection of the points $x^{k+j}$ is performed without the information on the trial results at the points $x^{k+i}$, $1 \leq i < j$.
+Only the first point $x^{k+1}$ will match to the point selected by the sequential algorithm.
+The points of other trials, in general, may not match the ones generated by the sequential algorithm.
+Therefore, we will call such trials redundant and the quantity
+\begin{displaymath}
+\lambda(p) = \left\{ \begin{array}{ll}
+                (n(p) - n(1)) / n(p), & \textrm{$n(p) > n(1)$}\\
+                0, & \textrm{$n(p) \leq n(1)$}
+  \end{array} \right.
+\end{displaymath}
+we will call the redundancy of the method.
+
+The following statements from \cite{Strongin2000} define the degree of parallelization $p$, which corresponds to the non-redundant (i.e. with zero redundancy) parallelization.
+
+Let us denote the series of trials generated by the sequential algorithm and by the parallel one when solving the same problem with $\varepsilon=0$ in the stopping conditions as $\{x^k\}$ and $\{y^m\}$, respectively.
+
+\textbf{Theorem}. Let $x^*$ be the global minimum point and $x^{\prime}$ be the local minimum one of the function $f(x)$ and let the following conditions be satisfied:
+    \begin{enumerate}
+        \item The inequality is fulfilled
+        $$ f(x^{\prime}) - f(x^*) \leq \delta, \delta > 0. \eqno(28)$$
+        \item The first $q(l)$ trials of the sequential method and of the parallel one match to each other i. e.
+        $$ \{x^1,...,x^{q(l)}\} = \{y^1,...,y^{q(l)}\} \eqno(29)$$
+        where
+         $$ \{x^1,...,x^{q(l)}\} \subset \{x^k\}, \{y^1,...,y^{q(l)}\}\subset \{y^k\}. \eqno(30)$$
+        \item There exist a point $t^n \in \{y^m\}$, $n < q(l)$ such as $x^{\prime} \leq y^n \leq x^*$ or $x^* \leq y^n \leq x^{\prime}$.
+        \item For the quantity $M$ from \cite{}, the inequality is fulfilled 
+        $$ M > 2^{2 - 1/N} H. \eqno(31)$$
+        where $H$ is Hölder constant of the minimized one-dimensional function.
+    \end{enumerate}
+Then, the parallel algorithm at $p=2$ will be non-redundant (i. e. $s(2)=2$, $\lambda(2)=0$) as long as the condition is fulfilled 
+$$ (x_{t_j} - x_{t_j - 1})^{1/N} > \frac{4\delta}{M - 2^{2 - 1/N} H}, j=1,2. \eqno(32)$$
+where $t_j$ is defined according to \cite{}.\\
+\textbf{Consequence}.
+Let the objective function $f(x)$ have $Q$ points of local minima $\{x_1^{\prime},...,x_Q^{\prime}\}$, which the condition (28) is fulfilled for, and let there exist the trial points $y^{n_i}$, $1 \leq i \leq Q$ such as
+$$ y^{n_i} \in \{y^1,...y^{q(l)}\}, \eqno(33)$$
+$$ \alpha_i \leq y^{n_i} \leq \alpha_{i+1}, \alpha_i, \alpha_{i+1} \in \{x^*, x_1^{\prime},...,x_Q^{\prime}\}, 1 \leq i \leq Q. \eqno(16)$$
+then, if the conditions of Theorem are fulfilled, the parallel algorithm with the parallelization degree $Q+1$ will be non-redundant (i.
+e.
+$s(Q+1)=Q+1, \lambda(Q+1)=0$) until the condition (32) is fulfilled.\\
+The consequence from Theorem plays a special role in solving the multidimen-sional problems reduced to the one-dimensional ones using the Peano-like evolvent $y(x)$.
+The evolvent $y(x)$, being an approximation to the Peano curve, has the effect of «splitting» of the global minimum point $y^* \in D$ into several (up to $2^N$) preimages in the interval [0,1].
+Applying the parallel algorithm to minimize the reduced one-dimensional function, one can expect a zero redundancy at the parallelization degree up to $2^N+1$.\\
+Now, let us consider the case when after the stage of the search for the rough approximation to the global solution, the stage of its local refinement is performed.
+Assume that the fraction of the search trials necessary for the local refinement of the solution makes $\alpha$ of total number of trials.
+Then, according to Amdahl's law, the overall speedup $S(p)$, which can be obtained by parallelization into $p$ processes, cannot exceed the quantity
+$$ S(p) = {\Big(\alpha + \frac{1-\alpha}{p}\Big)}^{-1}.$$
+This relation will be true if one assumes the non-redundant parallelization of the global search stage.
+The increase in the efficiency of the parallel algorithm will be limited by the fraction of trials at the stage of the local refinement of the solution.
+So far, to increase the efficiency of parallelization (especially in the case of a large number of processes employed), one needs to parallelize not only the global search but the local one as well.
+Two particular local methods applicable to the optimization problems with the «black box» type functions as well as the methods of parallelization of these ones will be considered in the next section.
+
+\section{Local optimization methods}
+\subsection{BFGS method}
+
+BFGS (Broyden-Fletcher-Goldfarb-Shanno) method (see [Nocedal, Wright]) belongs to the class of quasi-Newton methods using the values of the function $f(x^k)$ and of its gradient $\nabla f(x^k)$ to organize the search.
+In this method, the function f is assumed to have the properties of the quadratic model that corresponds to the existence of a symmetric matrix of the second derivatives (Hessian) of the function $\nabla^2 f(x^k)$.\\
+The basic idea for all quasi-Newton methods consists in the construction of an ap-proximation of the actual Hessian $\nabla^2 f(x^k)$ at every step of the algorithm using a special matrix $B^k$ followed by selection of the descent direction $d^k$ as a solution of the following system:
+$$ B^k d^k = -\nabla f(x^k) \leftrightarrow d^k = -H^k \nabla f(x^k),$$
+where $H^k={(B^k)}^{-1}$.\\
+After obtaining the direction, a step is performed from current point $x^k$:
+$$ x^{k+1}=x^k+ \alpha_k d^k,$$
+where $\alpha_k>0$ is the adjustable step length.\\
+Assuming $B^0=E$, the Hessian approximation is recalculated by the formula $B^{k+1}= B^k+ U^k$ where $U^k$ are the corrections represented in the form of a matrix of the rank 1.
+The small rank of the corrections $U^k$ is necessary for efficient procedure of calculating the inverse matrix $H^{k+1}=(B^{k+1})^{-1}=(B^k+ U^k)^{-1}$.\\
+The main requirement to forming the corrections $U^k$ of particular form is the secant equation $B^{k+1} (x^{k+1} - x^k )= \nabla f(x^{k+1})- \nabla f(x^k)$, which is performed for all quasi-Newton methods.
+For more precise determination of the matrix $B^{k+1}$, it is necessary to impose additional requirements on the Hessian approximation.
+The set of such requirements is determined by BFGS rule.
+Assume:
+$$ s_k = x^{k+1} - x^k,$$
+$$ y_k = \nabla f(x^{k+1})- \nabla f(x^k).$$
+Then, the BFGS recalculation scheme takes the following form:
+$$ B^{k+1} = B^k - \frac{B^k s_k s_k^T B^k}{s_k^T B^k s_k} + \frac{y^k (y^k)^{T}}{(y^k)^{T} s_k},$$
+$$ H^{k+1} = (E - \rho_k s_k (y^k)^{T}) H^k (E - \rho_k y^k s_k^T) + \rho_k s_k s_k^T,$$
+where $\rho_k = 1/((y^k)^{T} s_k)$.\\
+For the systems with limited memory, a modification of BFGS method – the lim-ited BFGS in the limited space (L-BFGS-B) is used.
+The modified method stores the history of $\aleph^k = ((s_{k-i}, y^{k-i}))_i^l$ of the last $l$ vectors $s_k$ and $y^k$.
+From the history, the approximate form of the matrix $H^k$ is restored.\\
+The algorithm of the BFGS method can be represented by a sequence of the fol-lowing commands:
+
+Step 0. Initialize the starting point $x^0$. Set the precision $\varepsilon > 0$;
+
+Step 1. Determine the initial approximation $H^0$ (eithe by the unitary matrix $E$ or by $\nabla^2 f$ at the point $x^0$;
+
+Step 2. Determine the search direction: $d^k = -H^k \nabla f^k$;
+
+Step 3. Compute $x^{k+1} = x^k + \alpha_k d^k$;
+
+Step 4. Determine the vectors $s_k$ and $y^k$;
+
+Step 5. Compute the Hessian $H^{k+1}$;
+
+Step 6. Stop if the condition $\|\nabla f^k\| < \varepsilon$ is fulfilled.
+
+\subsubsection{Hooke-Jeeves method}
+Hooke-Jeeves method (see, e.g., \cite{Himmelblau} ) is the first-order one, i.e.
+it needs only the function values for the work.
+In this method, the search for the minimum at every step is performed as a result of a shift along some sample direction (step by a sample), which is constructed and then corrected as a result of special trial coordinate translations called constructing a configuration.
+
+Constructing a configuration from a point $y$ is performed by a mapping of $y$ into a point $\bar{y} = F(y)$, where $F$ is an operator of constructing a configuration.
+It is constructed so that the direction $(\bar{y} - y)$ is that of the decreasing of the function $\varphi$ in the nearness of $y$.
+To describe the operator $F$, let us introduce the following notations: $e^i$ is the $i^{th}$ unit coordinate vector and $h$ is the parameter determining the magnitude of the coordinate translation.
+Then, the transition from $y$ to $\bar{y}$ (i.e.
+the algorithm of constricting the configuration $\bar{y} = F(y)$) is performed according to the following rules.
+
+Step 0. Set $\bar{y} = y$.
+
+Step 1. For $y$ from 1 to $N$ do:
+
+if $\varphi(\bar{y} + he^i) < \varphi(\bar{y})$ then set $\bar{y} = \bar{y} + he^i$ else if $\varphi(\bar{y} - he^i) < \varphi(y)$ then $\bar{y} = y - he^i.$
+Next, the step-by-step description of the whole method is given.
+
+Step 0.
+Set the starting point $t^0$, the parameter of the method $\varepsilon > 0$, the parameter of constructing the configuration $h\gg\varepsilon$, and the parameter of the increasing of the step $\alpha=2$.
+
+Step 1. Set $y^1  = t^0, k = 0$.
+
+Step 2. Construct a configuration $t^{k+1} = F(y^{k +1})$.
+
+Step 3.
+If $\varphi(t^{k+1}) < \varphi(t^{k})$ then $k:=k+1$ and go to Step 4, else if $h \leq \varepsilon$, stop the search, if $h > \varepsilon$ then further actions depend on the  way how the point $y^{k+1}$ was constructed: whether the configuration was constructed using the step by sample (in this case, $k > 0$) or it was constructed from the point $t^0$(in this case, $k=0$).
+If $k=0$, cut $h$ in half ($h=h/2$) and go to Step 1.
+If $k>0$, set $t^0 = t^k, k=0$ and also go to Step 1.
+
+Step 4. Perform the step by example $y^{k+1} = t^k + \alpha (t^k - t^{k-1})$ and go to Step 2.
+
+The meaning of the actions performed within the steps 2, 3, and 4 can be explained as follows. \\
+Step 4 was introduced in order to enable the method to increase the shift magnitude within a single iteration rapidly in the case when the point $t^k$ is located far away from the solution.
+At that, prior to make $y^{k+1}$ current point of the iteration, a configuration is constructed from the point $y^{k+1}$.
+By these means, in the case of obtaining a point with $\varphi(t^{k+1})< \varphi(t^k)$, the next step by sample, in general, will be performed in a changed direction (relative to previous one) that allows adapting the method to variations in the function relief.\\
+Finally, upon obtaining a value $\varphi(t^{k+1}) \geq \varphi(t^k)$ at Step 3, an attempt to run the algorithm again from the point $t^0=t^k$ (better than the one found) is undertaken.
+At that, if such an attempt hasn’t been undertaken yet, the parameter $h$ is not changed, but if it has been undertaken already, $h$ is cut in half first.
+
+\subsubsection{Parallelization of the local optimization methods}
+
+Parallel computing of many objective function values at once is the operation, which can be performed in parallel efficiently in the application of the optimization methods.
+The local optimization methods are the iteration ones, therefore, the parallelization can be organized within a single iteration only.
+
+For example, in Hooke-Jeeves method one can parallelize the construction of configuration i.
+e.
+parallelize the loop at the second step of the algorithm.
+In this loop, one may not compute the function values at the points $\bar{y} + he^i$ or $\bar{y} - he^i$ immedeatedly but may store the coordinates of the ones in the buffer, and upon accumulating P points compute the function values at these ones in parallel.
+Afterwards, the accumulated information is analyzed and the coordinates of the point $\bar{y}$ are changed.
+This operation is repeated until all $N$ coordinates $\bar{y}$ are computed.
+
+The main work of the quasi-Newton L-BFGS-B method in every iteration is spent for the computing of the function gradient.
+When solving the problems with the ``black box'' functions, the analytical calculation of the gradient is impossible.
+The numerical estimate of the gradient using the finite difference formulae implies multiple computing of the objective function values.
+
+So far, in the L-BFGS-B algorithm, one can parallelize the computing of gradient at the second step of the algorithm.
+To do so, one needs to store all points, which the computing of the function values for the numerical estimate of the gradient are necessary in into an intermediate buffer, and then compute the function values at the points stored earlier in parallel.
+At the final stage, the gradient vector is formed thus completing its estimate.
+
+The general scheme of organizing the parallel computations for all the local methods employed is presented in Fig.
+\#.
+The objective function values can be computed on a single node using shared memory (OpenMP), on a cluster with distributed memory (MPI), or on a graphical accelerator using CUDA.
+
+\subsubsection{Acknowledgements} This study was supported by the Russian Science Foundation, project No.\,21-11-00204.
 

### 2022-04-12 2022/Sensors/sensors.tex
fix some typos
fix some errors

@@
 \address[1]{%
-$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhny Novgorod, Russia; anton.shtanyuk@itmm.unn.ru (A.Sh.), sysoyev@vmk.unn.ru (A.S.)}
+$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhny Novgorod, Russia; anton.shtanyuk@itmm.unn.ru (A.Sh.), alexander.sysoyev@itmm.unn.ru (A.S.)}
 %$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
@@
 A well-known approach for accelerating the kNN algorithm is to utilize dimensionality reduction methods based on the use of space-filling curves.
-In this paper, we take this approach further and propose an algorithm that employs multiple space-filling curves and is faster (with comparable quality) compared to the kNN algorithm which uses kd-trees to determine the nearest neighbors.
+In this paper, we take this approach further and propose an algorithm that employs multiple space-filling curves and is faster (with comparable quality) compared to the kNN algorithm which uses kd-trees to determine the nearest neighbours.
 A specific method for constructing multiple Peano curves is outlined, and statements are given about the preservation of object proximity information in the course of dimensionality reduction.
@@
 \item calculate the distance from the object being classified to each of the objects in the training set;
-\item select k objects of the training set with the minimum distance to them;
+\item select $k$ objects of the training set with the minimum distance to them;
 \item the class of the object being classified is the class most often found among the $k$ nearest neighbours.
@@
 \begin{equation}
-\omega_i \leftrightarrow y = (y_{i1}, y_{i2}, ..., y_{iN})
+\omega_i \leftrightarrow y = (y_{i1}, y_{i2}, ..., y_{iN}).
 \end{equation}
@@
 \textbf{Remark}.
-The conditions of the theorem distinguish a specific neighborhood of the point $y^*$.
-This neighborhood comprises only such neighbours of this point which can be obtained by  shifting $y^*$ parallel to one of the coordinate axes by a distance not exceeding $2^{-p}$.
+The conditions of the theorem distinguish a specific neighbourhood of the point $y^*$.
+This neighbourhood comprises only such neighbours of this point which can be obtained by  shifting $y^*$ parallel to one of the coordinate axes by a distance not exceeding $2^{-p}$.
 By varying the value of $j, 1 \leq j \leq N$, under the conditions of the theorem, one can identify the nearest neighbours of point $y^*$ along $N$ coordinate directions.
@@
 \begin{equation}
-\omega_i \rightarrow \{ x_i^0, x_i^1, ..., x_i^L \}
+\omega_i \rightarrow \{ x_i^0, x_i^1, ..., x_i^L \},
 \end{equation} 
@@
 
-Several datasets were synthesized to examine the search time for neighbours. First, DS-Random-1 and DS-Random-2 sets consisting of 100000 and 10000 objects, respectively, were generated.
+Several datasets were synthesized to examine the search time for neighbours. First, DS-Random-1 and DS-Random-2 sets consisting of $100\;000$ and $10\;000$ objects, respectively, were generated.
 
@@
 \begin{itemize}
-\item \texttt{n\_samples} = 10000 (number of objects),
+\item \texttt{n\_samples} = $10\;000$ (number of objects),
 \item \texttt{n\_features} =  $N$ (dimensionality),
@@
 
-The number of objects of the first class is about 17000, the number of objects of the second class is 43000. The total number of objects examined is 60000.
+The number of objects of the first class is about $17\;000$, the number of objects of the second class is $43\;000$. The total number of objects examined is $60\;000$.
 

### 2022-04-13 2022/Sensors/kNN.tex
fix formulas numbering

--- 2022-04-12 2022/Sensors/sensors.tex
@@
 Previously, a feature vector $y = (y_1, y_2, ..., y_N) \in R^N$, corresponding to each object under consideration was defined, i.e.
-\begin{equation}
+\[
 \omega_i \leftrightarrow y = (y_{i1}, y_{i2}, ..., y_{iN}).
-\end{equation}
+\]
 
 Using the Peano curve, each object is assigned the value of $x_i$, i.e.
-\begin{equation}
+\[
 \omega_i \leftrightarrow x_i, \; y(x_i) = (y_{i1}, y_{i2}, ..., y_{iN}).
-\end{equation}
+\]
 
@@
 Thus, the closest object to the test object has the following property:
-\begin{equation}
+\[
 d_{min} = \min \{ d_i: 1 \leq i \leq S \}.
-\end{equation}
+\]
 
@@
 Data preparation consists of calculating the correspondence $\omega_i \rightarrow x_i$ for each input object, including the test object, and ordering the objects by the value of $x$.
-\begin{equation}
+\[
 x^1(\omega^1) \leq x^2(\omega^2) \leq ... \leq x^S(\omega^S).
-\end{equation}
+\]
 
@@
 It follows from formulae (\ref{hypercubeD0}) and (\ref{maps}) that if an interval $d(M,v)$ is mapped onto the subcube $D_0(M,v)$, then there exists a family of subcubes
-\begin{equation}
+\[
 D_l(M,v) = y^l(d(M,v)), 1 \leq l \leq L,
-\end{equation} 
+\]
 connected with the corresponding subintervals $d_l(M,v_l) \subset [0,1]$, where $d(M,v) = d_0(M,v_0), v_0 = v$, such that 
-\begin{equation}
+\[
 D_l(M,v) = y^l(d_l(M,v)), 1 \leq l \leq L.
-\end{equation}
+\]
 
 Since the hypercube $D$ from (\ref{D}) belongs to the common part of the family of hypercubes (\ref{hypercubes}) (the boundary of the hypercube $D$ is highlighted in Fig.~\ref{fig2}), then, by introducing an additional function
-\begin{equation}
+\[
 g_0(y) = \max \{ |y_i|-2^{-1}: \; 1 \leq i \leq N \},
-\end{equation} 
+\] 
 the original hypercube $D$ can be represented as
-\begin{equation}
+\[
 D = \{y^l(x): x \in [0,1], \; g_0(y^l(x)) \leq 0 \}, \; 0 \leq l \leq L,
-\end{equation} 
+\] 
 i.e. $g_0(y) \leq 0$, if $y \in D$, and $g_0(y) > 0$ if otherwise. Hence, any point $y \in D$ has its preimage $x^l \in [0,1]$ at every mapping $y^l(x), \; 0 \leq l \leq L$, i.e.
-\begin{equation}
+\[
 y = y^l(x^l), \; x^l \in [0, 1], \; 0 \leq l \leq L.
-\end{equation}
+\]
 
@@
 \textbf{Theorem 1.} Let $y^*$ be an arbitrary point from the domain $D$ belonging to the interval with endpoints $y^\prime, y^{\prime\prime} \in D$, which differ in the values of the only coordinate, and let
-\begin{equation}
+\[
 |y_j^\prime - y^{\prime\prime}_j| \leq 2^{-p}, \; y^\prime_i = y^{\prime\prime}_i = y_i^*, \; 1 \leq i \leq N, i \ne j,
-\end{equation} 
+\] 
 where $p$ is an integer, $1 \leq p \leq L-1$, and $j$ is the number of coordinate whose values for points $y^*, y^\prime, y^{\prime\prime}$ are different.
 Then there exists at least one correspondence $y^l(x), 0 \leq l \leq L$, and preimages $x^*, x^\prime, x^{\prime\prime} \in [0, 1]$ such that
-\begin{equation}
+\[
  y^* = y^l(x^*), y^\prime = y^l(x^\prime), y^{\prime\prime} = y^l(x^{\prime\prime}), \\
 \max \{ |x^\prime - x^*|, |x^{\prime\prime} - x^*|, |x^\prime - x^{\prime\prime}| \} \leq 2^{-pN}.
-\end{equation}
+\]
 
@@
 For every object $\omega_i$ a set of its preimages on different one-dimensional scales is assigned 
-\begin{equation}
+\[
 \omega_i \rightarrow \{ x_i^0, x_i^1, ..., x_i^L \},
-\end{equation} 
+\] 
 where $l, \; 0 \leq l \leq L$, is the number of the one-dimensional interval.
@@
 The degree of proximity of a test object $\omega^*$ to one of the known objects $\omega_i$ can be expressed in terms of the distance
-\begin{equation}
+\[
 d_i^l = |x_i^l - x^{*l}|
-\end{equation} 
+\]
 on the $l$-th scale.
@@
 If we find the minimum distance
-\begin{equation}
+\[
 d_{min} = \min_{1\leq i\leq n} d(\omega_i,\omega^*) ,
-\end{equation} 
+\]
 we get the nearest neighbour.

### 2022-04-13 2022/RSD/paper.tex
fix some formulaе and references

@@
 \renewcommand\UrlFont{\color{blue}\rmfamily}
-\usepackage{amsmath} %?????????!!!!!!!
+\usepackage{amsmath} 
+\usepackage{multirow} 
 %
@@
 Let us consider a problem of searching the global minimum of an $N$-dimensional function $\varphi(y)$ in a hyperinterval $D$
-\begin{equation}\label{problemN}
-\varphi(y^*) = \min{\{ \varphi(y):y \in D \}}
-\end{equation}
-\begin{equation}\label{D}
-D = \{ y \in R^{N}: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}.
+\begin{gather}
+\varphi(y^*) = \min{\{ \varphi(y):y \in D \}} \label{problemN} \\ 
+D = \{ y \in R^{N}: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}. \label{D}
+\end{gather}
+
+The main assumption, which the algorithm considered are based on is the fulfillment of Lipschitz condition for the function $\varphi(y)$, i.e.
+\begin{equation} \label{lip_ref}
+\mid \varphi(y_1) - \varphi(y_2) \mid \leq L\|y_1 - y_2\|, y_1, y_2 \in D, 0 < L < \infty.
 \end{equation}
 
-The main assumption, which the algorithm considered are based on is the fulfillment of Lipschitz condition for the function $\varphi(y)$, i.e.
-$$ \mid \varphi(y_1) - \varphi(y_2) \mid \leq L\|y_1 - y_2\|, y_1, y_2 \in D, 0 < L < \infty. \eqno(3)$$
 Note that the Lipschitz constant $L$ featuring the rate of variation of the function is not known \textit{a priori}.
@@
 This approach is based on the idea of the dimensionality reduction using Peano-Hilbert curve $y(x)$ mapping the interval $[0,1]$ of the real axis onto an $N$-dimensional cube
-$$ \{ y \in R^{N}: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N\} = \{y(x): 0 \leq x \leq 1 \}.    \eqno(4)$$
+\begin{equation} \label{n_dem_cube_ref}
+\{ y \in R^{N}: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N\} = \{y(x): 0 \leq x \leq 1 \}.
+\end{equation}
 
@@
 The use of the space-filling curves allows reducing the multidimensional problem \ref{problemN} to a one-dimensional problem 
-$$\varphi(y^*) = \varphi(y(x^*)) = \min{\{ \varphi(y(x)): x \in [0, 1] \}}. \eqno(5)$$
+\begin{equation} \label{problem_ref}
+\varphi(y^*) = \varphi(y(x^*)) = \min{\{ \varphi(y(x)): x \in [0, 1] \}}
+\end{equation}
 and applying the efficient one-dimensional optimization methods to solve this one.
@@
 If the function $\varphi(y)$ in the domain $D$ satisfies Lipschitz condition, the function $\varphi(y(x))$ in the interval $[0,1]$ will satisfy uniform H{\"o}lder condition
-$$\mid \varphi(y(x_1)) - \varphi(y(x_2)) \mid \leq H {\mid x_1 - x_2 \mid}^{1/N}, x_1, x_2 \in [0,1], \eqno(6)$$
+\begin{equation} \label{hölder_ref}
+\mid \varphi(y(x_1)) - \varphi(y(x_2)) \mid \leq H {\mid x_1 - x_2 \mid}^{1/N}, x_1, x_2 \in [0,1],
+\end{equation}
 where H{\"o}lder constant $H$ is related to Lipschitz constant $L$ by the relation $H = 2L\sqrt{N+3}$.
 Therefore, without liming the generality, one can consider the minimization of the one-dimensional function
-$$ f(x) = \varphi(y(x)), x \in [0,1], \eqno(7)$$
+\begin{equation} \label{minim_fun_ref}
+f(x) = \varphi(y(x)), x \in [0,1],
+\end{equation}
 satisfying H{\"o}lder condition.
@@
 Step 2. Calculate the magnitudes 
-$$ \mu = \max\limits_{1 \leq i \leq k} \frac{\mid z_i - z_{i-1} \mid}{\Delta_i}, $$
-
-\begin{displaymath}
+\begin{equation}\label{M}
+\mu = \max\limits_{1 \leq i \leq k} \frac{\mid z_i - z_{i-1} \mid}{\Delta_i}, \;\;
 M = \left\{ \begin{array}{ll}
                 r\mu, & \textrm{$\mu > 0$,}\\
-                1, & \textrm{$\mu = 0$}.
+                1, & \textrm{$\mu = 0$},
   \end{array} \right.
-\end{displaymath}
+\end{equation}
 where $z_i = f(y(x_i)),$  $\Delta_i = {(x_i - x_{i-1})}^{1/N}$, and $r > 1$ is a predefined parameter of the method.
 
-Step 3. For each interval $(x_{i-1}, x_i)$, $1 \leq i \leq k$ compute the characteristic according to the formula
-$$ R(i) = \Delta_i + \frac{(z_i - z_{i-1})^2}{M^2 \Delta_i} - 2 \frac{z_i + z_{i-1}}{M}, 1 \leq i \leq k. $$
+Step 3. For each interval $(x_{i-1}, x_i)$, $1 \leq i \leq k,$ compute the \textit{characteristic} according to the formula
+\begin{equation}\label{R}
+R(i) = \Delta_i + \frac{(z_i - z_{i-1})^2}{M^2 \Delta_i} - 2 \frac{z_i + z_{i-1}}{M}, 1 \leq i \leq k. 
+\end{equation}
 
-Step 4. Arrange the characteristics $R(i)$, $1 \leq i \leq k$ in the decreasing order
-$$ R(t_1) \geq R(t_2) \geq ... \geq R(t_{k-1}) \geq R(t_{k+1}) $$
+Step 4. Arrange the characteristics $R(i)$, $1 \leq i \leq k,$ in the decreasing order
+\begin{equation} \label{OrderedR}
+ R(t_1) \geq R(t_2) \geq ... \geq R(t_{k-1}) \geq R(t_{k+1}) 
+\end{equation}
 and select $P$ largest characteristics with the interval indices $t_j$, $1 \leq j \leq P$.
@@
 Step 5. Perform new trials at the points $x^{k+j}$, $1 \leq j \leq P$ computed by the formulae
-$$ x^{k+1} = \frac{x_{t_j} + x_{t_j - 1}}{2} - sign(z_{t_j} - z_{t_j - 1}) \frac{1}{2r} {\Bigg[\frac{\mid z_{t_j} - z_{t_j - 1} \mid}{\mu}\Bigg]}^{N}, 1 < t_j < k+1.$$
-Put the results of the trials performed into the information database of the algo-rithm.\\
+$$ x^{k+1} = \frac{x_{t_j} + x_{t_j - 1}}{2} - sign(z_{t_j} - z_{t_j - 1}) \frac{1}{2r} {\Bigg[\frac{\mid z_{t_j} - z_{t_j - 1} \mid}{\mu}\Bigg]}^{N}.$$
+Put the results of the trials performed into the information database of the algorithm.
 
-Step 6. Check the stopping criterion $\Delta_{t_j} \leq \varepsilon$ where $t_j$ from \cite{} and $\varepsilon > 0$ is the predefined precision.
-
+Step 6. Check the stopping criterion $\Delta_{t_j} \leq \varepsilon$, where $t_j$ are from Step 4 and $\varepsilon > 0$ is the predefined precision. 
 After completing the algorithm work, take as the estimate of the solution of the problem \ref{problemN} 
@@
 
-The theoretical substantiation of the algorithm presented here see in more details in \cite{}.
-Here, note briefly that the characteristics of the intervals used in the algorithm \cite{} can be considered as a quantitative measure how particular interval is promising from the viewpoint of finding the global minimum inside this one.
-The inequalities \cite{} arrange the intervals according to the characteristics on the ones, and the trials are performed in parallel in p most promising intervals.
+The theoretical substantiation of the algorithm presented here see in more details in \cite{Strongin2000}.
+Here, note briefly that the characteristics of the intervals (\ref{R}) used in the algorithm can be considered as a quantitative measure how particular interval is promising from the viewpoint of finding the global minimum inside this one.
+The inequalities (\ref{OrderedR}) arrange the intervals according to the characteristics on the ones, and the trials are performed in parallel in p most promising intervals.
 
-\subsubsection{Theoretical estimates of the speedup of the parallel algorithm}
-Let us describe (according to \cite{}) the theoretical properties of the parallel algorithm featuring its speedup.
+\subsection{Theoretical estimates of the speedup of the parallel algorithm}
+Let us describe (according to \cite{Strongin2000}) the theoretical properties of the parallel algorithm featuring its speedup.
 In the problems being solved, the time of executing a single search trial exceeds the time of processing its results essentially.
 Therefore, the speedup in the number of trials is the key characteristic of the efficiency of the parallel global search algorithm
-$$s(p) = k(1)\frac{p}{k(p)}, \eqno(26)$$
-where $k(1)$ is the number of trials executed by the sequential method and $k(p)$ is the number of trials executed by the parallel method employing p parallel processes.\\
+\begin{equation} \label{par_trl_ref}
+s(p) = \frac{k(1)p}{k(p)},
+\end{equation}
+where $k(1)$ is the number of trials executed by the sequential method and $k(p)$ is the number of trials executed by the parallel method employing $p$ parallel processes.
 
@@
 Indeed, the sequential algorithm has the full information obtained at the previous $k$ iterations when selecting the point $x^{k+1}$ of the next $(k+1)^{th}$ trial.
-The parallel algorithm selects not one but $p$ points $x^{k+j}$, $1 \leq j \leq p$ on the base of the same information at the $(k+1)^{th}$ iteration.
+The parallel algorithm selects not one but $p$ points $x^{k+j}$, $1 \leq j \leq p,$ on the base of the same information at the $(k+1)^{th}$ iteration.
 It means that the selection of the points $x^{k+j}$ is performed without the information on the trial results at the points $x^{k+i}$, $1 \leq i < j$.
@@
         \item The inequality is fulfilled
-        $$ f(x^{\prime}) - f(x^*) \leq \delta, \delta > 0. \eqno(28)$$
-        \item The first $q(l)$ trials of the sequential method and of the parallel one match to each other i. e.
-        $$ \{x^1,...,x^{q(l)}\} = \{y^1,...,y^{q(l)}\} \eqno(29)$$
+            \begin{equation} \label{first_s_ref}
+                f(x^{\prime}) - f(x^*) \leq \delta, \delta > 0.
+            \end{equation}
+        \item The first $q(l)$ trials of the sequential method and of the parallel one match to each other, i.e.
+            \begin{equation} \label{second_s_ref}
+                \{x^1,...,x^{q(l)}\} = \{y^1,...,y^{q(l)}\}
+            \end{equation}
         where
-         $$ \{x^1,...,x^{q(l)}\} \subset \{x^k\}, \{y^1,...,y^{q(l)}\}\subset \{y^k\}. \eqno(30)$$
+            \begin{equation} \label{third_s_ref}
+                \{x^1,...,x^{q(l)}\} \subset \{x^k\}, \{y^1,...,y^{q(l)}\}\subset \{y^k\}.
+            \end{equation}
         \item There exist a point $t^n \in \{y^m\}$, $n < q(l)$ such as $x^{\prime} \leq y^n \leq x^*$ or $x^* \leq y^n \leq x^{\prime}$.
-        \item For the quantity $M$ from \cite{}, the inequality is fulfilled 
-        $$ M > 2^{2 - 1/N} H. \eqno(31)$$
-        where $H$ is Hölder constant of the minimized one-dimensional function.
+        \item For the quantity $M$ from (\ref{M}), the inequality is fulfilled 
+            \begin{equation} \label{fourth_s_ref}
+                M > 2^{2 - 1/N} H.
+            \end{equation}
+        where $H$ is H{\"o}lder constant of the objective one-dimensional function.
     \end{enumerate}
-Then, the parallel algorithm at $p=2$ will be non-redundant (i. e. $s(2)=2$, $\lambda(2)=0$) as long as the condition is fulfilled 
-$$ (x_{t_j} - x_{t_j - 1})^{1/N} > \frac{4\delta}{M - 2^{2 - 1/N} H}, j=1,2. \eqno(32)$$
-where $t_j$ is defined according to \cite{}.\\
+
+Then, the parallel algorithm at $p=2$ will be non-redundant (i.e. $s(2)=2$, $\lambda(2)=0$) as long as the condition is fulfilled 
+\begin{equation} \label{p_two_ref}
+    (x_{t_j} - x_{t_j - 1})^{1/N} > \frac{4\delta}{M - 2^{2 - 1/N} H}, j=1,2.
+\end{equation}
+where $t_j$ is defined according to Step 4 of the algorithm.
+
 \textbf{Consequence}.
-Let the objective function $f(x)$ have $Q$ points of local minima $\{x_1^{\prime},...,x_Q^{\prime}\}$, which the condition (28) is fulfilled for, and let there exist the trial points $y^{n_i}$, $1 \leq i \leq Q$ such as
-$$ y^{n_i} \in \{y^1,...y^{q(l)}\}, \eqno(33)$$
-$$ \alpha_i \leq y^{n_i} \leq \alpha_{i+1}, \alpha_i, \alpha_{i+1} \in \{x^*, x_1^{\prime},...,x_Q^{\prime}\}, 1 \leq i \leq Q. \eqno(16)$$
-then, if the conditions of Theorem are fulfilled, the parallel algorithm with the parallelization degree $Q+1$ will be non-redundant (i.
-e.
-$s(Q+1)=Q+1, \lambda(Q+1)=0$) until the condition (32) is fulfilled.\\
+Let the objective function $f(x)$ have $Q$ points of local minima $\{x_1^{\prime},...,x_Q^{\prime}\}$, which the condition \eqref{first_s_ref} is fulfilled for, and let there exist the trial points $y^{n_i}$, $1 \leq i \leq Q$ such as
+\begin{gather} 
+    y^{n_i} \in \{y^1,...y^{q(l)}\}, \label{conseq_range_ref} \\ 
+    \alpha_i \leq y^{n_i} \leq \alpha_{i+1}, \alpha_i, \alpha_{i+1} \in \{x^*, x_1^{\prime},...,x_Q^{\prime}\}, 1 \leq i \leq Q. \label{conseq_ref}
+\end{gather}
+Then, if the conditions of theorem are fulfilled, the parallel algorithm with the parallelization degree $Q+1$ will be non-redundant (i.e.
+$s(Q+1)=Q+1, \lambda(Q+1)=0$) until the condition \eqref{p_two_ref} is fulfilled.
+
 The consequence from Theorem plays a special role in solving the multidimen-sional problems reduced to the one-dimensional ones using the Peano-like evolvent $y(x)$.
 The evolvent $y(x)$, being an approximation to the Peano curve, has the effect of «splitting» of the global minimum point $y^* \in D$ into several (up to $2^N$) preimages in the interval [0,1].
-Applying the parallel algorithm to minimize the reduced one-dimensional function, one can expect a zero redundancy at the parallelization degree up to $2^N+1$.\\
+Applying the parallel algorithm to minimize the reduced one-dimensional function, one can expect a zero redundancy at the parallelization degree up to $2^N+1$.
+
 Now, let us consider the case when after the stage of the search for the rough approximation to the global solution, the stage of its local refinement is performed.
@@
 $$ S(p) = {\Big(\alpha + \frac{1-\alpha}{p}\Big)}^{-1}.$$
+
 This relation will be true if one assumes the non-redundant parallelization of the global search stage.
@@
 So far, to increase the efficiency of parallelization (especially in the case of a large number of processes employed), one needs to parallelize not only the global search but the local one as well.
-Two particular local methods applicable to the optimization problems with the «black box» type functions as well as the methods of parallelization of these ones will be considered in the next section.
+Two particular local methods applicable to the optimization problems with the ``black-box'' type functions as well as the methods of parallelization of these ones will be considered in the next section.
 
-\section{Local optimization methods}
+\section{Local optimization methods}\label{sec:local}
 \subsection{BFGS method}
 
-BFGS (Broyden-Fletcher-Goldfarb-Shanno) method (see [Nocedal, Wright]) belongs to the class of quasi-Newton methods using the values of the function $f(x^k)$ and of its gradient $\nabla f(x^k)$ to organize the search.
-In this method, the function f is assumed to have the properties of the quadratic model that corresponds to the existence of a symmetric matrix of the second derivatives (Hessian) of the function $\nabla^2 f(x^k)$.\\
-The basic idea for all quasi-Newton methods consists in the construction of an ap-proximation of the actual Hessian $\nabla^2 f(x^k)$ at every step of the algorithm using a special matrix $B^k$ followed by selection of the descent direction $d^k$ as a solution of the following system:
+BFGS (Broyden-Fletcher-Goldfarb-Shanno) method (see \cite{Nocedal}) belongs to the class of quasi-Newton methods using the values of the function $f(x^k)$ and of its gradient $\nabla f(x^k)$ to organize the search.
+In this method, the function f is assumed to have the properties of the quadratic model that corresponds to the existence of a symmetric matrix of the second derivatives (Hessian) of the function $\nabla^2 f(x^k)$.
+
+The basic idea for all quasi-Newton methods consists in the construction of an approximation of the actual Hessian $\nabla^2 f(x^k)$ at every step of the algorithm using a special matrix $B^k$ followed by selection of the descent direction $d^k$ as a solution of the following system:
 $$ B^k d^k = -\nabla f(x^k) \leftrightarrow d^k = -H^k \nabla f(x^k),$$
-where $H^k={(B^k)}^{-1}$.\\
+where $H^k={(B^k)}^{-1}$.
+
 After obtaining the direction, a step is performed from current point $x^k$:
 $$ x^{k+1}=x^k+ \alpha_k d^k,$$
-where $\alpha_k>0$ is the adjustable step length.\\
+where $\alpha_k>0$ is the adjustable step length.
+
 Assuming $B^0=E$, the Hessian approximation is recalculated by the formula $B^{k+1}= B^k+ U^k$ where $U^k$ are the corrections represented in the form of a matrix of the rank 1.
-The small rank of the corrections $U^k$ is necessary for efficient procedure of calculating the inverse matrix $H^{k+1}=(B^{k+1})^{-1}=(B^k+ U^k)^{-1}$.\\
+The small rank of the corrections $U^k$ is necessary for efficient procedure of calculating the inverse matrix $H^{k+1}=(B^{k+1})^{-1}=(B^k+ U^k)^{-1}$.
+
 The main requirement to forming the corrections $U^k$ of particular form is the secant equation $B^{k+1} (x^{k+1} - x^k )= \nabla f(x^{k+1})- \nabla f(x^k)$, which is performed for all quasi-Newton methods.
@@
 $$ y_k = \nabla f(x^{k+1})- \nabla f(x^k).$$
+
 Then, the BFGS recalculation scheme takes the following form:
@@
 $$ H^{k+1} = (E - \rho_k s_k (y^k)^{T}) H^k (E - \rho_k y^k s_k^T) + \rho_k s_k s_k^T,$$
-where $\rho_k = 1/((y^k)^{T} s_k)$.\\
-For the systems with limited memory, a modification of BFGS method – the lim-ited BFGS in the limited space (L-BFGS-B) is used.
+where $\rho_k = 1/((y^k)^{T} s_k)$.
+
+For the systems with limited memory, a modification of BFGS method – the limited BFGS in the limited space (L-BFGS-B) is used.
 The modified method stores the history of $\aleph^k = ((s_{k-i}, y^{k-i}))_i^l$ of the last $l$ vectors $s_k$ and $y^k$.
-From the history, the approximate form of the matrix $H^k$ is restored.\\
+From the history, the approximate form of the matrix $H^k$ is restored.
+
 The algorithm of the BFGS method can be represented by a sequence of the fol-lowing commands:
@@
 
-Step 6. Stop if the condition $\|\nabla f^k\| < \varepsilon$ is fulfilled.
+Step 6. Stop, if the condition $\|\nabla f^k\| < \varepsilon$ is fulfilled.
 
-\subsubsection{Hooke-Jeeves method}
-Hooke-Jeeves method (see, e.g., \cite{Himmelblau} ) is the first-order one, i.e.
+\subsection{Hooke-Jeeves method}
+Hooke-Jeeves method (see, e.g., \cite{Himmelblau}) is the first-order one, i.e.
 it needs only the function values for the work.
@@
 if $\varphi(\bar{y} + he^i) < \varphi(\bar{y})$ then set $\bar{y} = \bar{y} + he^i$ else if $\varphi(\bar{y} - he^i) < \varphi(y)$ then $\bar{y} = y - he^i.$
+
 Next, the step-by-step description of the whole method is given.
@@
 
-The meaning of the actions performed within the steps 2, 3, and 4 can be explained as follows. \\
+The meaning of the actions performed within the steps 2, 3, and 4 can be explained as follows. 
 Step 4 was introduced in order to enable the method to increase the shift magnitude within a single iteration rapidly in the case when the point $t^k$ is located far away from the solution.
 At that, prior to make $y^{k+1}$ current point of the iteration, a configuration is constructed from the point $y^{k+1}$.
-By these means, in the case of obtaining a point with $\varphi(t^{k+1})< \varphi(t^k)$, the next step by sample, in general, will be performed in a changed direction (relative to previous one) that allows adapting the method to variations in the function relief.\\
+By these means, in the case of obtaining a point with $\varphi(t^{k+1})< \varphi(t^k)$, the next step by sample, in general, will be performed in a changed direction (relative to previous one) that allows adapting the method to variations in the function relief.
+
 Finally, upon obtaining a value $\varphi(t^{k+1}) \geq \varphi(t^k)$ at Step 3, an attempt to run the algorithm again from the point $t^0=t^k$ (better than the one found) is undertaken.
@@
 
-\subsubsection{Parallelization of the local optimization methods}
+\subsection{Parallelization of the local optimization methods}
 
@@
 parallelize the loop at the second step of the algorithm.
-In this loop, one may not compute the function values at the points $\bar{y} + he^i$ or $\bar{y} - he^i$ immedeatedly but may store the coordinates of the ones in the buffer, and upon accumulating P points compute the function values at these ones in parallel.
+In this loop, one may not compute the function values at the points $\bar{y} + he^i$ or $\bar{y} - he^i$ immedeatedly but may store the coordinates of the ones in the buffer, and upon accumulating $P$ points compute the function values at these ones in parallel.
 Afterwards, the accumulated information is analyzed and the coordinates of the point $\bar{y}$ are changed.
@@
 
+\section{Results of numerical experiments}
+The numerical experiments were carried out on Lobachevsky supercomputer.
+Each supercomputer node was a two-processor one (two Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM).
+All methods considered in the above sections were implemented using Globalizer software \cite{globalizerSystem} (development language – C++).
+Intel C++ 17.0.0 compiler was used to assembly the system.
+
+The first series of experiments was carried out with the local optimization algorithms only. Rosenbrock function is a classical test function for the local search methods 
+\begin{gather}
+    \varphi(y) = \sum_{i=1}^{N-1} \left((1-y_i)^2 + 100(y_{i+1} - y_i^2)^2\right) \nonumber \\
+    D = \{ y \in R^N: -25 \leq y_j \leq 25, 1 \leq j \leq N \}. \nonumber
+\end{gather}
+
+Tables XX and XX present the averaged numbers of iterations, which Hooke-Jeeves and BFGS methods executed for minimization of Rosenbrock function with the dimensionalities from 2 to 30.
+Tables XX and XX present corresponding speedups in iterations.
+The number of searches varied from 1 to 16.
+
+\begin{table}[ht]
+	\caption{Number of trials executed by Hooke-Jeeves method}
+	\label{tab:1}
+	\center
+	\begin{tabular}{|c|c|c|c|c|c|}
+		\hline
+		%\noalign{\smallskip}
+		N & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
+		\hline 
+		2 &	 &  &  &  &   \\
+		\hline
+		3 &	 &  &  &  &   \\
+		\hline
+		5 &	 &  &  &  &   \\
+		\hline
+		10 &  &  &  &  &   \\
+		\hline
+		15 &  &  &  &  &   \\
+		\hline
+		30 &  &  &  &  &   \\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+	\caption{Speedup when using Hooke-Jeeves method}
+	\label{tab:2}
+	\center
+	\begin{tabular}{|c|c|c|c|c|}
+		\hline
+		%\noalign{\smallskip}
+		N & P = 2 & P = 4 & P = 8 & P = 16 \\
+		\hline 
+		2 &  &  &  &   \\
+		\hline
+		3 &  &  &  &   \\
+		\hline
+		5 &  &  &  &   \\
+		\hline
+		10 &  &  &  &   \\
+		\hline
+		15 &  &  &  &   \\
+		\hline
+		30 &  &  &  &   \\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+	\caption{Number of trials executed by BFGS method}
+	\label{tab:3}
+	\center
+	\begin{tabular}{|c|c|c|c|c|c|}
+		\hline
+		%\noalign{\smallskip}
+		N & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
+		\hline 
+		2 &	 &  &  &  &   \\
+		\hline
+		3 &	 &  &  &  &   \\
+		\hline
+		5 &	 &  &  &  &   \\
+		\hline
+		10 &  &  &  &  &   \\
+		\hline
+		15 &  &  &  &  &   \\
+		\hline
+		30 &  &  &  &  &   \\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+	\caption{Speedup when using BFGS method}
+	\label{tab:4}
+	\center
+	\begin{tabular}{|c|c|c|c|c|}
+		\hline
+		%\noalign{\smallskip}
+		N & P = 2 & P = 4 & P = 8 & P = 16 \\
+		\hline 
+		2 &  &  &  &   \\
+		\hline
+		3 &  &  &  &   \\
+		\hline
+		5 &  &  &  &   \\
+		\hline
+		10 &  &  &  &   \\
+		\hline
+		15 &  &  &  &   \\
+		\hline
+		30 &  &  &  &   \\
+		\hline
+	\end{tabular}
+\end{table}
+
+The results demonstrated the methods of parallelization of the local optimization algorithms described in Section \ref{sec:local} to show good results for the test function.
+At that, for the problems with low dimensionality (less than 30) Hooke-Jeeves method suited better.
+However, starting from the dimensionality of 30, L-BFGS-B method begins to execute fewer trials.
+Also, because of some features of parallelization, Hooke-Jeeves method executes more and more computations of the objective function with in-creasing number of threads, whereas L-BFGS-B always performs the same number of trials.
+It is explained by the parallelization of the computing of gradient in L-BFGS-B so that the number of computations of the objective function is the same regardless to the number of threads while in Hooke-Jeeves method the algorithm itself is parallelized so that not all the points, at which the trials were executed earlier in the parallel algorithm match the trajectory of the sequential method.
+
+Next, the experiments, which the global search algorithms as well as the local methods have been employed in, were carried out.
+In \cite{Gaviano2003} GKLS (this abbreviation consists of the first letters of the authors’ names) generator was described.
+This generator allows generating the multiextremal optimization problems with the properties known \textit{a priori}: the number of local minima, the sizes of the attraction regions of the ones, the global minima points, the objective function value at this point, etc.
+The test problems generated by this generator are featured by a small time of computing the objective function values.
+
+The global minimum $y^*$ was considered to be found if the algorithm generates a trial point $y^k$ in the $\delta$-nearness of the global minimum, i.e.
+$\|y^k -y^*\| \leq \delta$.
+At that, the nearness size was selected (according to \cite{7}) as $\delta = \| b - a \|\sqrt[N]{\Delta}$.
+Here $N$ is the dimensionality of the problem being solved, $a$ and $b$ are the boundaries of the search domain $D$, the parameter $\Delta = 10^{-6}$ for $N = 4$ and $\Delta = 10^{-7}$ for $N = 5$.
+When using AGS method, the parameter $r=4.5$ was selected for class Simple, and $r=5.6$ for class Hard; the parameter of the construction of the Peano curve was fixed to be $m=10$.
+The maximum allowed number of iterations was $K_{max}= 5\cdot10^6$.
+
+The number of iterations for the local methods was limited to $10^4$, the precision of the local methods was set to $10^{-4}$.
+First, the problems were solved by the global search method with the parameters specified above.
+Then, a local method was started from the optimal point found to find the minimum more precisely.
+
+Table XX presents the averaged number of iterations, which AGS required to solve the problem.
+Tables XX and XX present the numbers of trials executed by Hooke-Jeeves and BFGS local methods.
+Tables XX and XX present the speedups achieved when solving the problem as a whole.
+
+\begin{table}[ht]
+	\caption{Averaged number of iterations for AGS algorithm}
+	\label{tab:5}
+	\center
+	\begin{tabular}{|c|c|c|c|c|c|c|}
+		\hline
+		N & Problem class & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
+		\hline 
+		    \multirow{2}{*}{4} & hard &  &  &  &  &  \\ \cline{2-7}
+		                       & simple &  &  &  &  &  \\
+		\hline
+		    \multirow{2}{*}{5} & hard &  &  &  &  &  \\ \cline{2-7}
+		                       & simple &  &  &  &  &  \\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+	\caption{Averaged number of trials executed by Hooke-Jeeves method}
+	\label{tab:6}
+	\center
+	\begin{tabular}{|c|c|c|c|c|c|c|}
+		\hline
+		N & Problem class & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
+		\hline 
+		    \multirow{2}{*}{4} & hard &  &  &  &  &  \\ \cline{2-7}
+		                       & simple &  &  &  &  &  \\
+		\hline
+		    \multirow{2}{*}{5} & hard &  &  &  &  &  \\ \cline{2-7}
+		                       & simple &  &  &  &  &  \\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+	\caption{Speedup when using Hooke-Jeeves method}
+	\label{tab:7}
+	\center
+	\begin{tabular}{|c|c|c|c|c|c|}
+		\hline
+		N & Problem class  & P = 2 & P = 4 & P = 8 & P = 16 \\
+		\hline 
+		    \multirow{2}{*}{4} & hard &  &  &  &  \\ \cline{2-6}
+		                       & simple &  &  &  &  \\
+		\hline
+		    \multirow{2}{*}{5} & hard &  &  &  &  \\ \cline{2-6}
+		                       & simple &  &  &  &  \\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+	\caption{Averaged number of trials executed by BFGS method}
+	\label{tab:8}
+	\center
+	\begin{tabular}{|c|c|c|c|c|c|c|}
+		\hline
+		N & Problem class & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
+		\hline 
+		    \multirow{2}{*}{4} & hard &  &  &  &  &  \\ \cline{2-7}
+		                       & simple &  &  &  &  &  \\
+		\hline
+		    \multirow{2}{*}{5} & hard &  &  &  &  &  \\ \cline{2-7}
+		                       & simple &  &  &  &  &  \\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+	\caption{Speedup when using BFGS method}
+	\label{tab:9}
+	\center
+	\begin{tabular}{|c|c|c|c|c|c|c|}
+		\hline
+		N & Problem class & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
+		\hline 
+		    \multirow{2}{*}{4} & hard &  &  &  &  &  \\ \cline{2-7}
+		                       & simple &  &  &  &  &  \\
+		\hline
+		    \multirow{2}{*}{5} & hard &  &  &  &  &  \\ \cline{2-7}
+		                       & simple &  &  &  &  &  \\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+	\caption{Speedup when using BFGS method}
+	\label{tab:10}
+	\center
+	\begin{tabular}{|c|c|c|c|c|c|}
+		\hline
+		N & Problem class  & P = 2 & P = 4 & P = 8 & P = 16 \\
+		\hline 
+		    \multirow{2}{*}{4} & hard &  &  &  &  \\ \cline{2-6}
+		                       & simple &  &  &  &  \\
+		\hline
+		    \multirow{2}{*}{5} & hard &  &  &  &  \\ \cline{2-6}
+		                       & simple &  &  &  &  \\
+		\hline
+	\end{tabular}
+\end{table}
+
+The local methods required less number of trials to refine the solution as compared to the global search method.
+As a result, the contributions of the local methods into total time of solving the problem were not very large.
+At that, L-BFGS-B method made much less computations of the objective function.
+
+Within the framework of the last experiment, an applied optimization problem was solved. This problem arose when solving the inverse problem of chemical kinetics.
+
+The reaction investigated and respective experimental data are described in \cite{Uskov2020}.
+The mathematical model of the chemical kinetics problem comprises a system of differential equations, which describes the variations of the concentrations of substances in time according to the rates of the reaction stages.
+According to this model, one can formulate an optimization problem, which defines the objective function as a sum of absolute deviations of the calculated concentrations from the ones measured experimentally:
+\[
+F = \sum_{i=1}^{M} \sum_{j=1}^{N} \mid x_{ij}^{calc} - x_{ij}^{exp}\mid \rightarrow \min,
+\]
+where $x_{ij}^{calc}$ and $x_{ij}^{exp}$ are the calculated values of concentrations of the components and the ones measured experimentally, respectively; $M$ is the number of measurement points, and $N$ is the number of substances participating in the reaction.
+The complete problem statement is given in \cite{PAVT}.
+Here, we note only that the number of the unknown parameters (i.e.
+the dimensionality of the optimization problem being solved) was 15, and we will consider this one as an optimization problem with a complex ``black-box'' objective function only.
+
+To solve the problem stated, the synchronous parallelization scheme (using MPI) was employed.
+The number of processes P to compute the function values varied from 1 up to 15.
+Also, there was an additional managing process, which processed the accumulated search information.
+In the stopping criteria of the method, the value $\varepsilon=0.01$ was used, the reliability parameter of the method was set as $r = 3$.
+After the work of GSA, a parallel local method was started to refine the result found, the precision of the local method was set to $10^{-4}$, the limitation on the number of trials was $10^4$.
+
+Table XX presents the total time of solving four problems corresponding to four different temperature regimes.
+\begin{table}[ht]
+	\caption{Averaged time of solving the problem, sec.}
+	\label{tab:11}
+	\center
+		\begin{tabular}{|c|c|c|c|c|}
+		\hline
+		%\noalign{\smallskip}
+		Method & P = 1 & P = 3 & P = 5 & P = 15 \\
+		\hline 
+		Hooke-Jeeves &  &  &  &   \\
+		\hline
+		L-BFGS-B &  &  &  &   \\
+		\hline
+	\end{tabular}
+\end{table}
+
+Table XX presents the data for one of the temperature regimes only (that corresponds to a single optimization problem): the time of solving without the run of a local method (Column GSA), using a sequential local method (Column GSA-l), and using a parallel local method run on 15 processes (GSA-pl).
+
+\begin{table}[ht]
+	\caption{Time of solving the problem using a local method}
+	\label{tab:12}
+	\center
+		\begin{tabular}{|c|c|c|c|}
+		\hline
+		%\noalign{\smallskip}
+		Method & GSA & GSA-l & GSA-pl \\
+		\hline 
+		Hooke-Jeeves &  &  &   \\
+		\hline
+		L-BFGS-B &  &  &   \\
+		\hline
+	\end{tabular}
+\end{table}
+
+Concluding general results, one can say that the use of local methods at the final stage of search allows solving the optimization problems with higher precision.
+In many cases, it is enough to use sequential local methods.
+However, for the problems with complex gully structure of the objective function require parallelizing the whole search process including the local stage as well.
+In the present work, two local optimization algorithms with the parallelization variants for shared memory as well as for distributed one are described.
+As the experiments have shown, the methods for parallelizing the local methods (at the final stage of global search) proposed demonstrated good results in solving both the test optimization problems and the applied ones.
+
 \subsubsection{Acknowledgements} This study was supported by the Russian Science Foundation, project No.\,21-11-00204.

### 2022-04-13 2022/RSD/paper.tex
Добавлены картинки.

@@
 
-\begin{table}[ht]
-	\caption{Number of trials executed by Hooke-Jeeves method}
-	\label{tab:1}
-	\center
-	\begin{tabular}{|c|c|c|c|c|c|}
-		\hline
-		%\noalign{\smallskip}
-		N & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
-		\hline 
-		2 &	 &  &  &  &   \\
-		\hline
-		3 &	 &  &  &  &   \\
-		\hline
-		5 &	 &  &  &  &   \\
-		\hline
-		10 &  &  &  &  &   \\
-		\hline
-		15 &  &  &  &  &   \\
-		\hline
-		30 &  &  &  &  &   \\
-		\hline
-	\end{tabular}
-\end{table}
+%\begin{table}[ht]
+%	\caption{Number of trials executed by Hooke-Jeeves method}
+%	\label{tab:1}
+%	\center
+%	\begin{tabular}{|c|c|c|c|c|c|}
+%		\hline
+%		%\noalign{\smallskip}
+%		N & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
+%		\hline 
+%		2 &	 &  &  &  &   \\
+%		\hline
+%		3 &	 &  &  &  &   \\
+%		\hline
+%		5 &	 &  &  &  &   \\
+%		\hline
+%		10 &  &  &  &  &   \\
+%		\hline
+%		15 &  &  &  &  &   \\
+%		\hline
+%		30 &  &  &  &  &   \\
+%		\hline
+%	\end{tabular}
+%\end{table}
+%
+%\begin{table}[ht]
+%	\caption{Speedup when using Hooke-Jeeves method}
+%	\label{tab:2}
+%	\center
+%	\begin{tabular}{|c|c|c|c|c|}
+%		\hline
+%		%\noalign{\smallskip}
+%		N & P = 2 & P = 4 & P = 8 & P = 16 \\
+%		\hline 
+%		2 &  &  &  &   \\
+%		\hline
+%		3 &  &  &  &   \\
+%		\hline
+%		5 &  &  &  &   \\
+%		\hline
+%		10 &  &  &  &   \\
+%		\hline
+%		15 &  &  &  &   \\
+%		\hline
+%		30 &  &  &  &   \\
+%		\hline
+%	\end{tabular}
+%\end{table}
+%
+%\begin{table}[ht]
+%	\caption{Number of trials executed by BFGS method}
+%	\label{tab:3}
+%	\center
+%	\begin{tabular}{|c|c|c|c|c|c|}
+%		\hline
+%		%\noalign{\smallskip}
+%		N & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
+%		\hline 
+%		2 &	 &  &  &  &   \\
+%		\hline
+%		3 &	 &  &  &  &   \\
+%		\hline
+%		5 &	 &  &  &  &   \\
+%		\hline
+%		10 &  &  &  &  &   \\
+%		\hline
+%		15 &  &  &  &  &   \\
+%		\hline
+%		30 &  &  &  &  &   \\
+%		\hline
+%	\end{tabular}
+%\end{table}
+%
+%\begin{table}[ht]
+%	\caption{Speedup when using BFGS method}
+%	\label{tab:4}
+%	\center
+%	\begin{tabular}{|c|c|c|c|c|}
+%		\hline
+%		%\noalign{\smallskip}
+%		N & P = 2 & P = 4 & P = 8 & P = 16 \\
+%		\hline 
+%		2 &  &  &  &   \\
+%		\hline
+%		3 &  &  &  &   \\
+%		\hline
+%		5 &  &  &  &   \\
+%		\hline
+%		10 &  &  &  &   \\
+%		\hline
+%		15 &  &  &  &   \\
+%		\hline
+%		30 &  &  &  &   \\
+%		\hline
+%	\end{tabular}
+%\end{table}
 
-\begin{table}[ht]
-	\caption{Speedup when using Hooke-Jeeves method}
-	\label{tab:2}
-	\center
-	\begin{tabular}{|c|c|c|c|c|}
-		\hline
-		%\noalign{\smallskip}
-		N & P = 2 & P = 4 & P = 8 & P = 16 \\
-		\hline 
-		2 &  &  &  &   \\
-		\hline
-		3 &  &  &  &   \\
-		\hline
-		5 &  &  &  &   \\
-		\hline
-		10 &  &  &  &   \\
-		\hline
-		15 &  &  &  &   \\
-		\hline
-		30 &  &  &  &   \\
-		\hline
-	\end{tabular}
-\end{table}
+\begin{figure}
+\begin{center}
+  \includegraphics[width=1.0\linewidth]{RosenbrocNT.jpg}
+  \caption{Number of trials executed on the Rosenbrock function}
+  \label{fig:RosenbrocNT}  
+\end{center}
+\end{figure}
 
-\begin{table}[ht]
-	\caption{Number of trials executed by BFGS method}
-	\label{tab:3}
-	\center
-	\begin{tabular}{|c|c|c|c|c|c|}
-		\hline
-		%\noalign{\smallskip}
-		N & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
-		\hline 
-		2 &	 &  &  &  &   \\
-		\hline
-		3 &	 &  &  &  &   \\
-		\hline
-		5 &	 &  &  &  &   \\
-		\hline
-		10 &  &  &  &  &   \\
-		\hline
-		15 &  &  &  &  &   \\
-		\hline
-		30 &  &  &  &  &   \\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[ht]
-	\caption{Speedup when using BFGS method}
-	\label{tab:4}
-	\center
-	\begin{tabular}{|c|c|c|c|c|}
-		\hline
-		%\noalign{\smallskip}
-		N & P = 2 & P = 4 & P = 8 & P = 16 \\
-		\hline 
-		2 &  &  &  &   \\
-		\hline
-		3 &  &  &  &   \\
-		\hline
-		5 &  &  &  &   \\
-		\hline
-		10 &  &  &  &   \\
-		\hline
-		15 &  &  &  &   \\
-		\hline
-		30 &  &  &  &   \\
-		\hline
-	\end{tabular}
-\end{table}
+\begin{figure}
+\begin{center}
+  \includegraphics[width=1.0\linewidth]{RosenbrocSpeedup.jpg}
+  \caption{Speedup on the Rosenbrock function}
+  \label{fig:RosenbrocSpeedup}  
+\end{center}
+\end{figure}
 
@@
 		\hline 
-		    \multirow{2}{*}{4} & hard &  &  &  &  &  \\ \cline{2-7}
-		                       & simple &  &  &  &  &  \\
+		    \multirow{2}{*}{4} & hard & 36528 & 18651 & 9854 & 4808 & 2260 \\ \cline{2-7}
+		                       & simple & 9995 & 5199 & 2706 & 1333 & 630 \\
 		\hline
-		    \multirow{2}{*}{5} & hard &  &  &  &  &  \\ \cline{2-7}
-		                       & simple &  &  &  &  &  \\
+		    \multirow{2}{*}{5} & hard & 266420 & 149109 & 90894 & 39873 & 22206 \\ \cline{2-7}
+		                       & simple & 29853 & 12913 & 7296 & 3618 & 2220 \\
 		\hline
@@
 
-\begin{table}[ht]
-	\caption{Averaged number of trials executed by Hooke-Jeeves method}
-	\label{tab:6}
-	\center
-	\begin{tabular}{|c|c|c|c|c|c|c|}
-		\hline
-		N & Problem class & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
-		\hline 
-		    \multirow{2}{*}{4} & hard &  &  &  &  &  \\ \cline{2-7}
-		                       & simple &  &  &  &  &  \\
-		\hline
-		    \multirow{2}{*}{5} & hard &  &  &  &  &  \\ \cline{2-7}
-		                       & simple &  &  &  &  &  \\
-		\hline
-	\end{tabular}
-\end{table}
+%\begin{table}[ht]
+%	\caption{Averaged number of trials executed by Hooke-Jeeves method}
+%	\label{tab:6}
+%	\center
+%	\begin{tabular}{|c|c|c|c|c|c|c|}
+%		\hline
+%		N & Problem class & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
+%		\hline 
+%		    \multirow{2}{*}{4} & hard &  &  &  &  &  \\ \cline{2-7}
+%		                       & simple &  &  &  &  &  \\
+%		\hline
+%		    \multirow{2}{*}{5} & hard &  &  &  &  &  \\ \cline{2-7}
+%		                       & simple &  &  &  &  &  \\
+%		\hline
+%	\end{tabular}
+%\end{table}
+%
+%\begin{table}[ht]
+%	\caption{Speedup when using Hooke-Jeeves method}
+%	\label{tab:7}
+%	\center
+%	\begin{tabular}{|c|c|c|c|c|c|}
+%		\hline
+%		N & Problem class  & P = 2 & P = 4 & P = 8 & P = 16 \\
+%		\hline 
+%		    \multirow{2}{*}{4} & hard &  &  &  &  \\ \cline{2-6}
+%		                       & simple &  &  &  &  \\
+%		\hline
+%		    \multirow{2}{*}{5} & hard &  &  &  &  \\ \cline{2-6}
+%		                       & simple &  &  &  &  \\
+%		\hline
+%	\end{tabular}
+%\end{table}
+%
+%
+%\begin{table}[ht]
+%	\caption{Averaged number of trials executed by BFGS method}
+%	\label{tab:8}
+%	\center
+%	\begin{tabular}{|c|c|c|c|c|c|c|}
+%		\hline
+%		N & Problem class & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
+%		\hline 
+%		    \multirow{2}{*}{4} & hard &  &  &  &  &  \\ \cline{2-7}
+%		                       & simple &  &  &  &  &  \\
+%		\hline
+%		    \multirow{2}{*}{5} & hard &  &  &  &  &  \\ \cline{2-7}
+%		                       & simple &  &  &  &  &  \\
+%		\hline
+%	\end{tabular}
+%\end{table}
+%
+%
+%\begin{table}[ht]
+%	\caption{Speedup when using BFGS method}
+%	\label{tab:10}
+%	\center
+%	\begin{tabular}{|c|c|c|c|c|c|}
+%		\hline
+%		N & Problem class  & P = 2 & P = 4 & P = 8 & P = 16 \\
+%		\hline 
+%		    \multirow{2}{*}{4} & hard &  &  &  &  \\ \cline{2-6}
+%		                       & simple &  &  &  &  \\
+%		\hline
+%		    \multirow{2}{*}{5} & hard &  &  &  &  \\ \cline{2-6}
+%		                       & simple &  &  &  &  \\
+%		\hline
+%	\end{tabular}
+%\end{table}
 
-\begin{table}[ht]
-	\caption{Speedup when using Hooke-Jeeves method}
-	\label{tab:7}
-	\center
-	\begin{tabular}{|c|c|c|c|c|c|}
-		\hline
-		N & Problem class  & P = 2 & P = 4 & P = 8 & P = 16 \\
-		\hline 
-		    \multirow{2}{*}{4} & hard &  &  &  &  \\ \cline{2-6}
-		                       & simple &  &  &  &  \\
-		\hline
-		    \multirow{2}{*}{5} & hard &  &  &  &  \\ \cline{2-6}
-		                       & simple &  &  &  &  \\
-		\hline
-	\end{tabular}
-\end{table}
+\begin{figure}
+\begin{center}
+  \includegraphics[width=1.0\linewidth]{GKLSNT.jpg}
+  \caption{Number of trials executed on the GKLS function}
+  \label{fig:GKLSNT}  
+\end{center}
+\end{figure}
 
-\begin{table}[ht]
-	\caption{Averaged number of trials executed by BFGS method}
-	\label{tab:8}
-	\center
-	\begin{tabular}{|c|c|c|c|c|c|c|}
-		\hline
-		N & Problem class & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
-		\hline 
-		    \multirow{2}{*}{4} & hard &  &  &  &  &  \\ \cline{2-7}
-		                       & simple &  &  &  &  &  \\
-		\hline
-		    \multirow{2}{*}{5} & hard &  &  &  &  &  \\ \cline{2-7}
-		                       & simple &  &  &  &  &  \\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[ht]
-	\caption{Speedup when using BFGS method}
-	\label{tab:9}
-	\center
-	\begin{tabular}{|c|c|c|c|c|c|c|}
-		\hline
-		N & Problem class & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
-		\hline 
-		    \multirow{2}{*}{4} & hard &  &  &  &  &  \\ \cline{2-7}
-		                       & simple &  &  &  &  &  \\
-		\hline
-		    \multirow{2}{*}{5} & hard &  &  &  &  &  \\ \cline{2-7}
-		                       & simple &  &  &  &  &  \\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[ht]
-	\caption{Speedup when using BFGS method}
-	\label{tab:10}
-	\center
-	\begin{tabular}{|c|c|c|c|c|c|}
-		\hline
-		N & Problem class  & P = 2 & P = 4 & P = 8 & P = 16 \\
-		\hline 
-		    \multirow{2}{*}{4} & hard &  &  &  &  \\ \cline{2-6}
-		                       & simple &  &  &  &  \\
-		\hline
-		    \multirow{2}{*}{5} & hard &  &  &  &  \\ \cline{2-6}
-		                       & simple &  &  &  &  \\
-		\hline
-	\end{tabular}
-\end{table}
+\begin{figure}
+\begin{center}
+  \includegraphics[width=1.0\linewidth]{GKLSSpeedup.jpg}
+  \caption{Speedup on the GKLS function}
+  \label{fig:GKLSSpeedup}  
+\end{center}
+\end{figure}
 
@@
 Table XX presents the total time of solving four problems corresponding to four different temperature regimes.
+
 \begin{table}[ht]
@@
 	\center
-		\begin{tabular}{|c|c|c|c|c|}
+		\begin{tabular}{|c|c|c|c|}
 		\hline
 		%\noalign{\smallskip}
-		Method & P = 1 & P = 3 & P = 5 & P = 15 \\
+		Method & P = 1 & P = 3 &  P = 15 \\
 		\hline 
-		Hooke-Jeeves &  &  &  &   \\
+		Hooke-Jeeves & 2386.4 & 1053.1 & 713.1  \\
 		\hline
-		L-BFGS-B &  &  &  &   \\
+		L-BFGS-B & 2571.7  & 1096.4  & 720.3  \\
 		\hline
@@
 		\hline 
-		Hooke-Jeeves &  &  &   \\
+		Hooke-Jeeves & 158.1 & 299.8  & 177.4 \\
 		\hline
-		L-BFGS-B &  &  &   \\
+		L-BFGS-B & 158.1  & 252.5  & 165.6  \\
 		\hline

### 2022-04-13 2022/RSD/paper.tex
Sections 1-3 proofreaded

--- 2022-04-13 2022/RSD/paper.tex
@@
 Two stages of solving the problem were distinguished: the search of a rough approximation to the global solution and its local refining.
-The novelty of the parallel algorithm considered in the present pa-per consists in the parallelization of the whole computational process – both at the global search stage and at the one of the local refining of the solution found.
+The novelty of the parallel algorithm considered in the present pa-per consists in the parallelization of the whole computational process -- both at the global search stage and at the one of the local refining of the solution found.
 The theoretical estimates of the algorithm parallelization efficiency are presented.
@@
 
-As a rule, the objective function $\varphi(y)$ in such problems is a ``black-box''-one, i.e.
+As a rule, the objective function $\varphi(y)$ in such problems is a ``black-box'' function, i.e.
 it is defined not by a formula but as a software-implemented algorithm for computing its values at the points of the search domain $D$.
-In this case, the analytical methods for the search of the exact problem solution is $y^{*} \in D$, and its numerical solving is reduced to the construction of an approximate solution $y_k^* \in D, \|y^{*} - y_k^*\| \leq \varepsilon$, where $\varepsilon > 0$ is a predefined precision.
-Hereafter, the process of computing the objective function at some point will be called a search trial.
+In this case, the analytical methods for the search of the exact problem solution $y^{*}  \in D$ are not applicable, and its numerical solving is reduced to the construction of an approximate solution $y_k^* \in D, \|y^{*} - y_k^*\| \leq \varepsilon$, where $\varepsilon > 0$ is a predefined accuracy.
+Hereafter, the process of computing the objective function at some point will be called a \textit{search trial}.
 
@@
 The solution of a global optimization problem always implies performing the search trials at the nodes of some grid (uniform, random, or adaptive one) in the search domain.
-The number of nodes in the grid increases exponentially with increasing dimensionality of the problem being solved.
-At the same time, to find the local extremum, the function values are computed in the nodes constituting not a grid but some trajectory starting from the initial guess point and ending at the solution point.
+The number of the grid nodes increases exponentially with increasing dimensionality of the problem being solved.
+At the same time, to find the local extremum, the function values are computed in the nodes constituting not a grid but some trajectory starting from the initial  point and ending at the solution point.
 
 Formally, the point $y^{\prime} \in D$ is called the local minimizer for a function on the set $D$ if there exists such a number $\varepsilon > 0$ that for all $y \in D$ such that $\|y - y^{\prime}\| < \varepsilon$, the inequality $\varphi(y^{\prime}) \leq \varphi(y)$ holds.
-Since one needs to explore not the whole search domain $D$ to find a local minimizer, and, on the other hand, the descend trajectory is, generally speaking, a one-dimension object, the computational costs for the local optimization methods are smaller essentially than the ones for the global search.
+Since one needs to explore not the whole search domain $D$ to find a local minimizer, and the descend trajectory is, generally speaking, a one-dimension object, the computational costs for the local optimization methods are essentially smaller than the ones for the global search.
 One can say that the local minimum search problem consists in finding the local minimizer $y^{*}(y^{0})$ for given initial guess point $y^{0} \in D$, which falls into the attraction region of $y^{*}$.
 
-There are many cases requiring the search for the local optimal solutions.
-One of such cases arises when an estimate of the globally-optimal solution $y_k^*$ is obtained with a non-satisfactory precision at some $k^{th}$ iteration of the global optimization method.
-In this case, one can find a locally–optimal solution $y_{loc}^*$, (corresponding to the initial guess point $y_k^*$) with a high precision.
+There are many cases requiring the search for the local optimum.
+One of such cases arises when an estimate of the global solution $y_k^*$ is obtained with a non-satisfactory accuracy at some $k^{th}$ iteration of the global optimization method.
+In this case one can find a local optimizer $y_{loc}^*$, which corresponds to the starting point $y_k^*$, with a high accuracy.
 The local minimum $y_{loc}^*$ found depends, generally speaking, on $y_k^*$, i.e.
@@
 
-As a rule, the computational costs for the local refining of the solution is smaller essentially than the costs, which would be necessary for the multiextremal optimization procedures to achieve the same precision as the local method can ensure.
-How-ever, in the case of using the parallel global optimization algorithm (at the first stage) and sequential local method (at the second stage), the work times of these ones can become comparable.
-However, the parallelization efficiency would be low.
+As a rule, the computational costs for the local refining of the solution is essentially smaller than the costs, which would be necessary for the multiextremal optimization procedures to achieve the same precision as the local method can ensure.
+However, in the case of using the parallel global optimization algorithm (at the first stage) and sequential local method (at the second stage) the work times of these ones can become comparable and the parallelization efficiency would be low.
 
-The novelty of the parallel algorithm considered in the present paper consists in the parallelization of the whole computing process – both at the global search stage and at the one of the local refining of the solution found.
-This approach can be relevant for the problems, which the objective functions are gully-wise with a weak multiextremality.
-In such problems, the stage of searching the rough approximation of the global solution and the one of its local refining are comparable in the computation costs.
+The novelty of the parallel algorithm considered in the present paper consists in the parallelization of the whole computing process -- both at the global search stage and at the stage of the local refining of the solution found.
+This approach can be relevant for the problems whose objective functions are flat valleys with a weak multiextremality.
+In such problems, the phase of searching the rough approximation of the global solution and the phase of its local refining are comparable in the computation costs.
 For example, in the work \cite{Gubaydullin2021} the global search stage only was parallelized; the local refining was made with a sequential algorithm.
@@
 \begin{gather}
-\varphi(y^*) = \min{\{ \varphi(y):y \in D \}} \label{problemN} \\ 
+\varphi(y^*) = \min{\{ \varphi(y):y \in D \}}, \label{problemN} \\ 
 D = \{ y \in R^{N}: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}. \label{D}
@@
 
-The main assumption, which the algorithm considered are based on is the fulfillment of Lipschitz condition for the function $\varphi(y)$, i.e.
+The main assumption, which the algorithm considered are based on, is the fulfillment of Lipschitz condition for the function $\varphi(y)$, i.e.
 \begin{equation} \label{lip_ref}
-\mid \varphi(y_1) - \varphi(y_2) \mid \leq L\|y_1 - y_2\|, y_1, y_2 \in D, 0 < L < \infty.
+\left| \varphi(y_1) - \varphi(y_2) \right| \leq L\|y_1 - y_2\|, \; y_1, y_2 \in D, \; 0 < L < \infty.
 \end{equation}
@@
 
-In the theory, the Peano-Hilbert curve is a limiting object.
+In the theory, the Peano-Hilbert curve is a limit object.
 Therefore, one needs to construct an approximation to the theoretical curve for the practical application.
@@
 
-The use of the space-filling curves allows reducing the multidimensional problem \ref{problemN} to a one-dimensional problem 
+The use of the space-filling curves allows reducing the multidimensional problem \ref{problemN} to a one-dimensional problem and applying the efficient one-dimensional optimization methods to solve it, i.e.
 \begin{equation} \label{problem_ref}
-\varphi(y^*) = \varphi(y(x^*)) = \min{\{ \varphi(y(x)): x \in [0, 1] \}}
+\varphi(y^*) = \varphi(y(x^*)) = \min{\{ \varphi(y(x)): x \in [0, 1] \}}.
 \end{equation}
-and applying the efficient one-dimensional optimization methods to solve this one.
-Keeping the property of limited relative differences of the objective function is an important property of using the mappings of this kind.
-If the function $\varphi(y)$ in the domain $D$ satisfies Lipschitz condition, the function $\varphi(y(x))$ in the interval $[0,1]$ will satisfy uniform H{\"o}lder condition
-\begin{equation} \label{hölder_ref}
-\mid \varphi(y(x_1)) - \varphi(y(x_2)) \mid \leq H {\mid x_1 - x_2 \mid}^{1/N}, x_1, x_2 \in [0,1],
+Keeping the property of limited relative differences of the objective function is an important property of using the evolvents.
+If the function $\varphi(y)$ in the domain $D$ satisfies the Lipschitz condition, the function $\varphi(y(x))$ in the interval $[0,1]$ will satisfy the uniform H{\"o}lder condition
+\begin{equation} \label{holder_ref}
+\left| \varphi(y(x_1)) - \varphi(y(x_2)) \right| \leq H {\left| x_1 - x_2 \right|}^{1/N}, \; x_1, x_2 \in [0,1],
 \end{equation}
@@
 
-The algorithm of solving this problem being considered implies the construction of a series of points $x_k$, which the values of the objective function  $z_k = f(x_k)$ are computed in.
+The algorithm of solving this problem being considered implies the construction of a series of points $x^k$, which the values of the objective function  $f(x^k) = \varphi(y(x^k))$ are computed in.
 We will divide the solving of the problem into two stages.
-At the first stage, the search of a rough approximation to the global minimum point will be per-formed using the global search algorithm.
+At the first stage, the search of a rough approximation to the global minimum point will be performed using the global search algorithm.
 At the second stage, a local refining of the solution found will be conducted with a high precision.
-The methods applied at the first stage and at the second one are described in Section 3 and Section 4, respectively.
+The methods applied at the first stage and at the second one are described in Section \ref{sec:global} and Section \ref{sec:local}, respectively.
 
-\section{Parallel global search algorithm}
+\section{Parallel global search algorithm}\label{sec:global}
 \subsection{Computational scheme of the algorithm}
 
-We will consider the class of problems, in which the time of executing a single search trial doesn’t depend on the point of the trial execution.
-In this case, one can use the synchronous parallelization scheme, which implies $P$ trials to be performed in parallel and synchronously within single method iteration.
-In this case, total number of trials executed after $n$ parallel iterations will be $k=nP$.
+We will consider the class of problems, in which the time of executing a single search trial doesn’t depend on the trial point.
+In this case, one can use the synchronous parallelization scheme, which implies $p$ trials to be carried out in parallel and synchronously within single method iteration.
+In this case, total number of trials executed after $n$ parallel iterations will be $k=n \cdot p$.
 
 The first iteration of the method is performed in a special way.
-At this one, the trials are performed at the boundary points $x^0 = 0$ and $x^1 = 1$ as well as at $(P-2)$ arbitrary different internal points $x^2, x^3, ..., x^P$ in the interval $[0,1]$.
+The trials are performed at the boundary points $x^0 = 0$ and $x^1 = 1$ as well as at $(p-2)$ arbitrary different internal points $x^2, x^3, ..., x^{p-1}$ in the interval $[0,1]$.
 
@@
 $$ 0 = x_0 < x_1 < ... < x_k = 1.$$
+
 Step 2. Calculate the magnitudes 
@@
 \end{equation}
-where $z_i = f(y(x_i)),$  $\Delta_i = {(x_i - x_{i-1})}^{1/N}$, and $r > 1$ is a predefined parameter of the method.
+where $z_i = f(y(x_i)),$  $\Delta_i = {(x_i - x_{i-1})}^{1/N}$, and $r > 1$ is a predefined number (reliability parameter of the method).
 
@@
 \end{equation}
-and select $P$ largest characteristics with the interval indices $t_j$, $1 \leq j \leq P$.
+and select $p$ largest characteristics with the interval indices $t_j$, $1 \leq j \leq p$.
 
-Step 5. Perform new trials at the points $x^{k+j}$, $1 \leq j \leq P$ computed by the formulae
-$$ x^{k+1} = \frac{x_{t_j} + x_{t_j - 1}}{2} - sign(z_{t_j} - z_{t_j - 1}) \frac{1}{2r} {\Bigg[\frac{\mid z_{t_j} - z_{t_j - 1} \mid}{\mu}\Bigg]}^{N}.$$
+Step 5. Carry out new trials in parallel at the points $x^{k+j}$, $1 \leq j \leq p$, computed by the formulae
+$$ x^{k+j} = \frac{x_{t_j} + x_{t_j - 1}}{2} - sign(z_{t_j} - z_{t_j - 1}) \frac{1}{2r} {\Bigg[\frac{\left| z_{t_j} - z_{t_j - 1} \right|}{\mu}\Bigg]}^{N}.$$
 Put the results of the trials performed into the information database of the algorithm.
 
-Step 6. Check the stopping criterion $\Delta_{t_j} \leq \varepsilon$, where $t_j$ are from Step 4 and $\varepsilon > 0$ is the predefined precision. 
+Step 6. Check the stopping criterion $\Delta_{t_j} \leq \varepsilon$, where $t_j$ are from Step 4 and $\varepsilon > 0$ is the predefined accuracy. 
 After completing the algorithm work, take as the estimate of the solution of the problem \ref{problemN} 
-$$ f_k^* = \min\limits_{1 \leq i \leq k} f(x^i), x_k^* = \arg \min\limits_{1 \leq i \leq k} f(x^i).$$
+$$ f_k^* = \min\limits_{1 \leq i \leq k} f(x^i), \; x_k^* = \arg \min\limits_{1 \leq i \leq k} f(x^i).$$
 
 The theoretical substantiation of the algorithm presented here see in more details in \cite{Strongin2000}.
-Here, note briefly that the characteristics of the intervals (\ref{R}) used in the algorithm can be considered as a quantitative measure how particular interval is promising from the viewpoint of finding the global minimum inside this one.
-The inequalities (\ref{OrderedR}) arrange the intervals according to the characteristics on the ones, and the trials are performed in parallel in p most promising intervals.
+Here, note briefly that the characteristics of the intervals (\ref{R}) used in the algorithm can be considered as a quantitative measure how particular interval is promising from the viewpoint of finding the global minimum inside it.
+The inequalities (\ref{OrderedR}) arrange the intervals according to their characteristics, and the trials are performed in parallel in $p$ most promising intervals.
 
 \subsection{Theoretical estimates of the speedup of the parallel algorithm}
+
 Let us describe (according to \cite{Strongin2000}) the theoretical properties of the parallel algorithm featuring its speedup.
-In the problems being solved, the time of executing a single search trial exceeds the time of processing its results essentially.
-Therefore, the speedup in the number of trials is the key characteristic of the efficiency of the parallel global search algorithm
+In the problems being solved, the time of executing a single search trial essentially exceeds the time of processing its result.
+Therefore, the speedup in the number of trials
 \begin{equation} \label{par_trl_ref}
-s(p) = \frac{k(1)p}{k(p)},
+s(p) = \frac{n(1)p}{n(p)}
 \end{equation}
-where $k(1)$ is the number of trials executed by the sequential method and $k(p)$ is the number of trials executed by the parallel method employing $p$ parallel processes.
+is the key characteristic of the efficiency of the parallel global search algorithm.
+Here $n(1)$ is the number of trials executed by the sequential method and $n(p)$ is the number of trials executed by the parallel method employing $p$ parallel processes.
 
@@
 Indeed, the sequential algorithm has the full information obtained at the previous $k$ iterations when selecting the point $x^{k+1}$ of the next $(k+1)^{th}$ trial.
-The parallel algorithm selects not one but $p$ points $x^{k+j}$, $1 \leq j \leq p,$ on the base of the same information at the $(k+1)^{th}$ iteration.
+The parallel algorithm selects not one but $p$ points $x^{k+j}$, $1 \leq j \leq p,$ on the base of the same information.
 It means that the selection of the points $x^{k+j}$ is performed without the information on the trial results at the points $x^{k+i}$, $1 \leq i < j$.
@@
             \begin{equation} \label{third_s_ref}
-                \{x^1,...,x^{q(l)}\} \subset \{x^k\}, \{y^1,...,y^{q(l)}\}\subset \{y^k\}.
+                \{x^1,...,x^{q(l)}\} \subset \{x^k\}, \{y^1,...,y^{q(l)}\}\subset \{y^m\}.
             \end{equation}
-        \item There exist a point $t^n \in \{y^m\}$, $n < q(l)$ such as $x^{\prime} \leq y^n \leq x^*$ or $x^* \leq y^n \leq x^{\prime}$.
+        \item There exist a point $t^n \in \{y^m\}$, $n < q(l)$, such as $x^{\prime} \leq y^n \leq x^*$ or $x^* \leq y^n \leq x^{\prime}$.
         \item For the quantity $M$ from (\ref{M}), the inequality is fulfilled 
             \begin{equation} \label{fourth_s_ref}
-                M > 2^{2 - 1/N} H.
+                M > 2^{2 - 1/N} H,
             \end{equation}
@@
 \begin{equation} \label{p_two_ref}
-    (x_{t_j} - x_{t_j - 1})^{1/N} > \frac{4\delta}{M - 2^{2 - 1/N} H}, j=1,2.
+    (x_{t_j} - x_{t_j - 1})^{1/N} > \frac{4\delta}{M - 2^{2 - 1/N} H},
 \end{equation}
@@
 \begin{gather} 
-    y^{n_i} \in \{y^1,...y^{q(l)}\}, \label{conseq_range_ref} \\ 
-    \alpha_i \leq y^{n_i} \leq \alpha_{i+1}, \alpha_i, \alpha_{i+1} \in \{x^*, x_1^{\prime},...,x_Q^{\prime}\}, 1 \leq i \leq Q. \label{conseq_ref}
+    y^{n_i} \in \{y^1,...y^{q(l)}\}, \nonumber \\ 
+    \alpha_i \leq y^{n_i} \leq \alpha_{i+1}, \; \alpha_i, \alpha_{i+1} \in \{x^*, x_1^{\prime},...,x_Q^{\prime}\}, 1 \leq i \leq Q. \nonumber
 \end{gather}
 Then, if the conditions of theorem are fulfilled, the parallel algorithm with the parallelization degree $Q+1$ will be non-redundant (i.e.
-$s(Q+1)=Q+1, \lambda(Q+1)=0$) until the condition \eqref{p_two_ref} is fulfilled.
+$s(Q+1)=Q+1, \; \lambda(Q+1)=0$) until the condition \eqref{p_two_ref} is fulfilled.
 
-The consequence from Theorem plays a special role in solving the multidimen-sional problems reduced to the one-dimensional ones using the Peano-like evolvent $y(x)$.
-The evolvent $y(x)$, being an approximation to the Peano curve, has the effect of «splitting» of the global minimum point $y^* \in D$ into several (up to $2^N$) preimages in the interval [0,1].
+The consequence from theorem plays a special role in solving the multidimensional problems reduced to the one-dimensional ones using the evolvent $y(x)$.
+The evolvent $y(x)$, being an approximation to the Peano curve, has the effect of ``splitting'' of the global minimum point $y^* \in D$ into several (up to $2^N$) preimages in the interval [0,1].
 Applying the parallel algorithm to minimize the reduced one-dimensional function, one can expect a zero redundancy at the parallelization degree up to $2^N+1$.
@@
 parallelize the loop at the second step of the algorithm.
-In this loop, one may not compute the function values at the points $\bar{y} + he^i$ or $\bar{y} - he^i$ immedeatedly but may store the coordinates of the ones in the buffer, and upon accumulating $P$ points compute the function values at these ones in parallel.
+In this loop, one may not compute the function values at the points $\bar{y} + he^i$ or $\bar{y} - he^i$ immedeatedly but may store the coordinates of the ones in the buffer, and upon accumulating $p$ points compute the function values at these ones in parallel.
 Afterwards, the accumulated information is analyzed and the coordinates of the point $\bar{y}$ are changed.
@@
 To solve the problem stated, the synchronous parallelization scheme (using MPI) was employed.
-The number of processes P to compute the function values varied from 1 up to 15.
+The number of processes $p$ to compute the function values varied from 1 up to 15.
 Also, there was an additional managing process, which processed the accumulated search information.

### 2022-04-14 2022/RSD/paper.tex
first proofread of Local methods section

@@
 
-BFGS (Broyden-Fletcher-Goldfarb-Shanno) method (see \cite{Nocedal}) belongs to the class of quasi-Newton methods using the values of the function $f(x^k)$ and of its gradient $\nabla f(x^k)$ to organize the search.
-In this method, the function f is assumed to have the properties of the quadratic model that corresponds to the existence of a symmetric matrix of the second derivatives (Hessian) of the function $\nabla^2 f(x^k)$.
+BFGS (Broyden-Fletcher-Goldfarb-Shanno) method (see \cite{Nocedal}) belongs to the class of quasi-Newton methods using the values of the function $\varphi(y^k)$ and of its gradient $\nabla \varphi(y^k)$ to organize the search.
+In this method, the function $\varphi(y)$ is assumed to have the properties of the quadratic function that corresponds to the existence of a symmetric matrix of the second derivatives (Hessian) $\nabla^2 \varphi(y)$.
 
-The basic idea for all quasi-Newton methods consists in the construction of an approximation of the actual Hessian $\nabla^2 f(x^k)$ at every step of the algorithm using a special matrix $B^k$ followed by selection of the descent direction $d^k$ as a solution of the following system:
-$$ B^k d^k = -\nabla f(x^k) \leftrightarrow d^k = -H^k \nabla f(x^k),$$
-where $H^k={(B^k)}^{-1}$.
+The basic idea for all quasi-Newton methods consists in the construction of an approximation of the actual Hessian $\nabla^2 \varphi(y^k)$ at every step of the algorithm using a special matrix $B^k$ followed by selection of the descent direction $d^k$ as a solution of the system
+$$ B^k d^k = -\nabla \varphi(y^k) \leftrightarrow d^k = -H^k \nabla \varphi(y^k),$$
+where $H^k=\left(B^k\right)^{-1}$.
 
-After obtaining the direction, a step is performed from current point $x^k$:
-$$ x^{k+1}=x^k+ \alpha_k d^k,$$
+After obtaining the direction, a step is performed from current point $y^k$:
+$$ y^{k+1}=y^k+ \alpha_k d^k,$$
 where $\alpha_k>0$ is the adjustable step length.
 
-Assuming $B^0=E$, the Hessian approximation is recalculated by the formula $B^{k+1}= B^k+ U^k$ where $U^k$ are the corrections represented in the form of a matrix of the rank 1.
+Assuming $B^0=E$, the Hessian approximation is recalculated by the formula $B^{k+1}= B^k+ U^k$, where $U^k$ are the corrections represented in the form of a matrix of the rank 1.
 The small rank of the corrections $U^k$ is necessary for efficient procedure of calculating the inverse matrix $H^{k+1}=(B^{k+1})^{-1}=(B^k+ U^k)^{-1}$.
 
-The main requirement to forming the corrections $U^k$ of particular form is the secant equation $B^{k+1} (x^{k+1} - x^k )= \nabla f(x^{k+1})- \nabla f(x^k)$, which is performed for all quasi-Newton methods.
+The main requirement to forming the corrections $U^k$ of particular form is the secant equation $B^{k+1} (y^{k+1} - y^k )= \nabla \varphi(y^{k+1})- \nabla \varphi(y^k)$, which is performed for all quasi-Newton methods.
 For more precise determination of the matrix $B^{k+1}$, it is necessary to impose additional requirements on the Hessian approximation.
 The set of such requirements is determined by BFGS rule.
-Assume:
-$$ s_k = x^{k+1} - x^k,$$
-$$ y_k = \nabla f(x^{k+1})- \nabla f(x^k).$$
+Assume
+\begin{gather}
+s_k = y^{k+1} - y^k, \nonumber \\
+z^k = \nabla \varphi(y^{k+1})- \nabla \varphi(y^k). \nonumber
+\end{gather}
 
 Then, the BFGS recalculation scheme takes the following form:
-$$ B^{k+1} = B^k - \frac{B^k s_k s_k^T B^k}{s_k^T B^k s_k} + \frac{y^k (y^k)^{T}}{(y^k)^{T} s_k},$$
-$$ H^{k+1} = (E - \rho_k s_k (y^k)^{T}) H^k (E - \rho_k y^k s_k^T) + \rho_k s_k s_k^T,$$
+$$ B^{k+1} = B^k - \frac{B^k s_k s_k^T B^k}{s_k^T B^k s_k} + \frac{z^k (z^k)^{T}}{(z^k)^{T} s_k},$$
+$$ H^{k+1} = (E - \rho_k s_k (z^k)^{T}) H^k (E - \rho_k z^k s_k^T) + \rho_k s_k s_k^T,$$
 where $\rho_k = 1/((y^k)^{T} s_k)$.
 
-For the systems with limited memory, a modification of BFGS method – the limited BFGS in the limited space (L-BFGS-B) is used.
-The modified method stores the history of $\aleph^k = ((s_{k-i}, y^{k-i}))_i^l$ of the last $l$ vectors $s_k$ and $y^k$.
+For the systems with limited memory, a modification of BFGS method -- the limited BFGS in the limited space (L-BFGS-B) is used.
+The modified method stores the history $\aleph^k = ((s_{k-i}, z^{k-i}))_i^l$ of the last $l$ vectors $s_k$ and $z^k$.
 From the history, the approximate form of the matrix $H^k$ is restored.
 
-The algorithm of the BFGS method can be represented by a sequence of the fol-lowing commands:
+The algorithm of the BFGS method can be represented by a sequence of the following steps.
 
-Step 0. Initialize the starting point $x^0$. Set the precision $\varepsilon > 0$;
+Step 0. Initialize the starting point $y^0$. Set the precision $\varepsilon > 0$;
 
-Step 1. Determine the initial approximation $H^0$ (eithe by the unitary matrix $E$ or by $\nabla^2 f$ at the point $x^0$;
+Step 1. Determine the initial approximation $H^0$ (either by the unit matrix $E$ or by $\nabla^2 \varphi(y^0)$;
 
-Step 2. Determine the search direction: $d^k = -H^k \nabla f^k$;
+Step 2. Determine the search direction: $d^k = -H^k \nabla \varphi(y^k)$;
 
-Step 3. Compute $x^{k+1} = x^k + \alpha_k d^k$;
+Step 3. Compute $y^{k+1} = y^k + \alpha_k d^k$;
 
-Step 4. Determine the vectors $s_k$ and $y^k$;
+Step 4. Determine the vectors $s_k$ and $z^k$;
 
@@
 
-Step 6. Stop, if the condition $\|\nabla f^k\| < \varepsilon$ is fulfilled.
+Step 6. Stop, if the condition $\|\nabla \varphi(y^k)\| < \varepsilon$ is fulfilled.
 
@@
 it needs only the function values for the work.
-In this method, the search for the minimum at every step is performed as a result of a shift along some sample direction (step by a sample), which is constructed and then corrected as a result of special trial coordinate translations called constructing a configuration.
+In this method, the search for the minimum at every step is performed as a result of a shift along some sample direction (step by a sample), which is constructed and then corrected as a result of special trial coordinate translations called \textit{constructing a configuration}.
 
@@
 Step 0.
-Set the starting point $t^0$, the parameter of the method $\varepsilon > 0$, the parameter of constructing the configuration $h\gg\varepsilon$, and the parameter of the increasing of the step $\alpha=2$.
+Set the starting point $t^0$, the parameter of the method $\varepsilon > 0$, the parameter of constructing the configuration $h\gg\varepsilon$, and the parameter $\alpha=2$ of the increasing of the step, e.g.
+$\alpha=2$.
 
@@
 
-Step 4. Perform the step by example $y^{k+1} = t^k + \alpha (t^k - t^{k-1})$ and go to Step 2.
+Step 4. Perform the step by direction $y^{k+1} = t^k + \alpha (t^k - t^{k-1})$ and go to Step 2.
 
@@
 
-Parallel computing of many objective function values at once is the operation, which can be performed in parallel efficiently in the application of the optimization methods.
+Parallel computing of many objective function values at once is the operation, which can be performed efficiently in the optimization methods.
 The local optimization methods are the iteration ones, therefore, the parallelization can be organized within a single iteration only.
 
-For example, in Hooke-Jeeves method one can parallelize the construction of configuration i.
-e.
+For example, in Hooke-Jeeves method one can parallelize the construction of configuration, i.e.
 parallelize the loop at the second step of the algorithm.
-In this loop, one may not compute the function values at the points $\bar{y} + he^i$ or $\bar{y} - he^i$ immedeatedly but may store the coordinates of the ones in the buffer, and upon accumulating $p$ points compute the function values at these ones in parallel.
-Afterwards, the accumulated information is analyzed and the coordinates of the point $\bar{y}$ are changed.
+In this loop, one may not compute the function values at the points $\bar{y} + he^i$ or $\bar{y} - he^i$ immediately but may store the coordinates of the ones in the buffer, and upon accumulating $p$ points compute the function values at these ones in parallel.
+Next, the accumulated information is analyzed and the coordinates of the point $\bar{y}$ are changed.
 This operation is repeated until all $N$ coordinates $\bar{y}$ are computed.
 
-The main work of the quasi-Newton L-BFGS-B method in every iteration is spent for the computing of the function gradient.
+The main work of the quasi-Newton L-BFGS-B method at every iteration is spent for the computing of the function gradient.
 When solving the problems with the ``black box'' functions, the analytical calculation of the gradient is impossible.
@@
 So far, in the L-BFGS-B algorithm, one can parallelize the computing of gradient at the second step of the algorithm.
-To do so, one needs to store all points, which the computing of the function values for the numerical estimate of the gradient are necessary in into an intermediate buffer, and then compute the function values at the points stored earlier in parallel.
-At the final stage, the gradient vector is formed thus completing its estimate.
+To do so, one needs to store all points, in which the computing of the function values for the gradient estimate are required, into an intermediate buffer, and then compute the function values at these points in parallel.
+%At the final stage, the gradient vector is formed thus completing its estimate.
 
-The general scheme of organizing the parallel computations for all the local methods employed is presented in Fig.
+The general scheme of the parallel computing for all the local methods employed is presented in Fig.
 \#.
-The objective function values can be computed on a single node using shared memory (OpenMP), on a cluster with distributed memory (MPI), or on a graphical accelerator using CUDA.
+The objective function values can be computed on a single node using shared memory (OpenMP), on a cluster with distributed memory (MPI), or on accelerator using CUDA.
 

### 2022-04-14 2022/RSD/paper.tex
Добавлены картинки в pdf, изменена последняя таблица.
proofread Results section

@@
 
-We will consider the class of problems, in which the time of executing a single search trial doesn’t depend on the trial point.
+We will consider the class of problems, in which the time of executing a single search trial doesn't depend on the trial point.
 In this case, one can use the synchronous parallelization scheme, which implies $p$ trials to be carried out in parallel and synchronously within single method iteration.
@@
 
-The theoretical substantiation of the algorithm presented here see in more details in \cite{Strongin2000}.
-Here, note briefly that the characteristics of the intervals (\ref{R}) used in the algorithm can be considered as a quantitative measure how particular interval is promising from the viewpoint of finding the global minimum inside it.
+The theoretical substantiation of the presented algorithm (denoted hereafter as \textit{global search algorithm} or GSA) see in more details in \cite{Strongin2000}.
+Note briefly that the characteristics of the intervals (\ref{R}) used in the algorithm can be considered as a quantitative measure how particular interval is promising from the viewpoint of finding the global minimum inside it.
 The inequalities (\ref{OrderedR}) arrange the intervals according to their characteristics, and the trials are performed in parallel in $p$ most promising intervals.
@@
 
-BFGS (Broyden-Fletcher-Goldfarb-Shanno) method (see \cite{Nocedal}) belongs to the class of quasi-Newton methods using the values of the function $\varphi(y^k)$ and of its gradient $\nabla \varphi(y^k)$ to organize the search.
+BFGS (Broyden-Fletcher-Goldfarb-Shanno) method (see, e.g., \cite{Nocedal}) belongs to the class of quasi-Newton methods using the values of the function $\varphi(y^k)$ and of its gradient $\nabla \varphi(y^k)$ to organize the search.
 In this method, the function $\varphi(y)$ is assumed to have the properties of the quadratic function that corresponds to the existence of a symmetric matrix of the second derivatives (Hessian) $\nabla^2 \varphi(y)$.
@@
 \subsection{Hooke-Jeeves method}
-Hooke-Jeeves method (see, e.g., \cite{Himmelblau}) is the first-order one, i.e.
+Hooke-Jeeves method (see, e.g., \cite{Kelley}) is the first-order one, i.e.
 it needs only the function values for the work.
@@
 \section{Results of numerical experiments}
+
 The numerical experiments were carried out on Lobachevsky supercomputer.
@@
 
-Tables XX and XX present the averaged numbers of iterations, which Hooke-Jeeves and BFGS methods executed for minimization of Rosenbrock function with the dimensionalities from 2 to 30.
-Tables XX and XX present corresponding speedups in iterations.
-The number of searches varied from 1 to 16.
+Fig.
+XX presents the averaged numbers of iterations, which Hooke-Jeeves and BFGS methods executed for minimization of Rosenbrock function with the dimensionalities from 2 to 30.
+Fig.
+XX presents corresponding speedups in iterations.
+The number of parallel trials $p$ varied from 1 to 16.
 
@@
 However, starting from the dimensionality of 30, L-BFGS-B method begins to execute fewer trials.
-Also, because of some features of parallelization, Hooke-Jeeves method executes more and more computations of the objective function with in-creasing number of threads, whereas L-BFGS-B always performs the same number of trials.
+Also, because of some features of parallelization, Hooke-Jeeves method executes more and more trials with increasing number of threads, whereas L-BFGS-B always performs the same number of trials.
 It is explained by the parallelization of the computing of gradient in L-BFGS-B so that the number of computations of the objective function is the same regardless to the number of threads while in Hooke-Jeeves method the algorithm itself is parallelized so that not all the points, at which the trials were executed earlier in the parallel algorithm match the trajectory of the sequential method.
@@
 Next, the experiments, which the global search algorithms as well as the local methods have been employed in, were carried out.
-In \cite{Gaviano2003} GKLS (this abbreviation consists of the first letters of the authors’ names) generator was described.
+In \cite{Gaviano2003} GKLS (this abbreviation consists of the first letters of the authors' names) generator was described.
 This generator allows generating the multiextremal optimization problems with the properties known \textit{a priori}: the number of local minima, the sizes of the attraction regions of the ones, the global minima points, the objective function value at this point, etc.
@@
 
-The global minimum $y^*$ was considered to be found if the algorithm generates a trial point $y^k$ in the $\delta$-nearness of the global minimum, i.e.
-$\|y^k -y^*\| \leq \delta$.
-At that, the nearness size was selected (according to \cite{7}) as $\delta = \| b - a \|\sqrt[N]{\Delta}$.
+The global minimum $y^*$ was considered to be found if the algorithm generates a trial point $y^k$ near the global minimum, i.e. $\|y^k -y^*\| \leq \delta$. 
+We use $\delta = 0.06$ for $N=4$ and $\delta = 0.08$ for $N=5$.
+%At that, the nearness size was selected (according to \cite{7}) as $\delta = \| b - a \|\sqrt[N]{\Delta}$.
 Here $N$ is the dimensionality of the problem being solved, $a$ and $b$ are the boundaries of the search domain $D$, the parameter $\Delta = 10^{-6}$ for $N = 4$ and $\Delta = 10^{-7}$ for $N = 5$.
-When using AGS method, the parameter $r=4.5$ was selected for class Simple, and $r=5.6$ for class Hard; the parameter of the construction of the Peano curve was fixed to be $m=10$.
+When using GSA, the parameter $r=4.5$ was selected for class Simple, and $r=5.6$ for class Hard; the parameter of the construction of the Peano curve was fixed to be $m=10$.
 The maximum allowed number of iterations was $K_{max}= 5\cdot10^6$.
@@
 
-Table XX presents the averaged number of iterations, which AGS required to solve the problem.
-Tables XX and XX present the numbers of trials executed by Hooke-Jeeves and BFGS local methods.
-Tables XX and XX present the speedups achieved when solving the problem as a whole.
+Table XX presents the averaged number of iterations, which GSA required to solve the problem.
+Fig.
+presents the numbers of trials executed by Hooke-Jeeves and L-BFGS-B local methods.
+Fig.
+XX presents the speedups achieved when solving the problem as a whole.
 
 \begin{table}[ht]
-	\caption{Averaged number of iterations for AGS algorithm}
+	\caption{Averaged number of iterations for GSA algorithm}
 	\label{tab:5}
@@
 
-Within the framework of the last experiment, an applied optimization problem was solved. This problem arose when solving the inverse problem of chemical kinetics.
+Within the last experiment an applied optimization problem was solved. This problem arose when solving the inverse problem of chemical kinetics.
 
@@
 \[
-F = \sum_{i=1}^{M} \sum_{j=1}^{N} \mid x_{ij}^{calc} - x_{ij}^{exp}\mid \rightarrow \min,
+F = \sum_{i=1}^{M} \sum_{j=1}^{N} \left| x_{ij}^{calc} - x_{ij}^{exp} \right| \rightarrow \min,
 \]
 where $x_{ij}^{calc}$ and $x_{ij}^{exp}$ are the calculated values of concentrations of the components and the ones measured experimentally, respectively; $M$ is the number of measurement points, and $N$ is the number of substances participating in the reaction.
-The complete problem statement is given in \cite{PAVT}.
-Here, we note only that the number of the unknown parameters (i.e.
+The complete problem statement is given in \cite{Uskov2020}.
+Here we note only that the number of the unknown parameters (i.e.
 the dimensionality of the optimization problem being solved) was 15, and we will consider this one as an optimization problem with a complex ``black-box'' objective function only.
@@
 Also, there was an additional managing process, which processed the accumulated search information.
-In the stopping criteria of the method, the value $\varepsilon=0.01$ was used, the reliability parameter of the method was set as $r = 3$.
-After the work of GSA, a parallel local method was started to refine the result found, the precision of the local method was set to $10^{-4}$, the limitation on the number of trials was $10^4$.
+In the stopping criteria of the method the value $\varepsilon=0.01$ was used and the reliability parameter of the method was set as $r = 3$.
+After the work of GSA, a parallel local method was started to refine the result found.
+The precision of the local method was set to $10^{-4}$ and the limitation on the number of trials was $10^4$.
 
-Table XX presents the total time of solving four problems corresponding to four different temperature regimes.
+%Table XX presents the total time of solving four problems corresponding to four different temperature regimes.
+%
+%\begin{table}[ht]
+	%\caption{Averaged time of solving the problem, sec.}
+	%\label{tab:11}
+	%\center
+		%\begin{tabular}{|c|c|c|c|}
+		%\hline
+		%%\noalign{\smallskip}
+		%Method & P = 1 & P = 3 &  P = 15 \\
+		%\hline 
+		%Hooke-Jeeves & 2386.4 & 1053.1 & 713.1  \\
+		%\hline
+		%L-BFGS-B & 2571.7  & 1096.4  & 720.3  \\
+		%\hline
+	%\end{tabular}
+%\end{table}
 
-\begin{table}[ht]
-	\caption{Averaged time of solving the problem, sec.}
-	\label{tab:11}
-	\center
-		\begin{tabular}{|c|c|c|c|}
-		\hline
-		%\noalign{\smallskip}
-		Method & P = 1 & P = 3 &  P = 15 \\
-		\hline 
-		Hooke-Jeeves & 2386.4 & 1053.1 & 713.1  \\
-		\hline
-		L-BFGS-B & 2571.7  & 1096.4  & 720.3  \\
-		\hline
-	\end{tabular}
-\end{table}
-
-Table XX presents the data for one of the temperature regimes only (that corresponds to a single optimization problem): the time of solving without the run of a local method (Column GSA), using a sequential local method (Column GSA-l), and using a parallel local method run on 15 processes (GSA-pl).
+Table \ref{tab:12} presents the results of solving the optimization problem: the running time of parallel GSA without the run of a local method (column GSA), with a sequential local method (column GSA-l) and with a parallel local method run on 15 processes (column GSA-pl).
 

### 2022-04-14 2022/RSD/paper.tex

@@
 The novelty of the parallel algorithm considered in the present paper consists in the parallelization of the whole computing process -- both at the global search stage and at the stage of the local refining of the solution found.
-This approach can be relevant for the problems whose objective functions are flat valleys with a weak multiextremality.
+This approach can be relevant for the problems whose objective functions are ravine-type with a weak multiextremality.
 In such problems, the phase of searching the rough approximation of the global solution and the phase of its local refining are comparable in the computation costs.
@@
 
-Fig.
-XX presents the averaged numbers of iterations, which Hooke-Jeeves and BFGS methods executed for minimization of Rosenbrock function with the dimensionalities from 2 to 30.
-Fig.
-XX presents corresponding speedups in iterations.
+Fig.~\ref{fig:RosenbrocNT} presents the averaged numbers of iterations, which Hooke-Jeeves and BFGS methods executed for minimization of Rosenbrock function with the dimensionalities from 2 to 30.
+Fig.~\ref{fig:RosenbrocSpeedup} presents corresponding speedups in iterations.
 The number of parallel trials $p$ varied from 1 to 16.
@@
 \begin{center}
-  \includegraphics[width=1.0\linewidth]{RosenbrocNT.jpg}
-  \caption{Number of trials executed on the Rosenbrock function}
+  \includegraphics[width=1.0\linewidth]{RosenbrocNT.pdf}
+  \caption{Number of trials (in thousands) executed on the Rosenbrock function}
   \label{fig:RosenbrocNT}  
@@
 \begin{center}
-  \includegraphics[width=1.0\linewidth]{RosenbrocSpeedup.jpg}
+  \includegraphics[width=1.0\linewidth]{RosenbrocSpeedup.pdf}
   \caption{Speedup on the Rosenbrock function}
@@
 
-Table XX presents the averaged number of iterations, which GSA required to solve the problem.
-Fig.
-presents the numbers of trials executed by Hooke-Jeeves and L-BFGS-B local methods.
-Fig.
-XX presents the speedups achieved when solving the problem as a whole.
+Table~\ref{tab:5} presents the averaged number of iterations, which GSA required to solve the problem.
+Fig.~\ref{fig:GKLSNT} presents the numbers of trials executed by Hooke-Jeeves and L-BFGS-B local methods.
+Fig.~\ref{fig:GKLSSpeedup} presents the speedups achieved when solving the problem as a whole.
 
@@
 \begin{center}
-  \includegraphics[width=1.0\linewidth]{GKLSNT.jpg}
+  \includegraphics[width=1.0\linewidth]{GKLSNT.pdf}
   \caption{Number of trials executed on the GKLS function}
@@
 \begin{center}
-  \includegraphics[width=1.0\linewidth]{GKLSSpeedup.jpg}
+  \includegraphics[width=1.0\linewidth]{GKLSSpeedup.pdf}
   \caption{Speedup on the GKLS function}
@@
 To solve the problem stated, the synchronous parallelization scheme (using MPI) was employed.
-The number of processes $p$ to compute the function values varied from 1 up to 15.
+The number of processes to compute the function values was $p=15$.
 Also, there was an additional managing process, which processed the accumulated search information.
@@
 	%\end{tabular}
+%\begin{table}[ht]
+%	\caption{Averaged time of solving the problem, sec.}
+%	\label{tab:11}
+%	\center
+%		\begin{tabular}{|c|c|c|c|}
+%		\hline
+%		%\noalign{\smallskip}
+%		Method & P = 1 & P = 3 &  P = 15 \\
+%		\hline 
+%		Hooke-Jeeves & 2386.4 & 1053.1 & 713.1  \\
+%		\hline
+%		L-BFGS-B & 2571.7  & 1096.4  & 720.3  \\
+%		\hline
+%	\end{tabular}
 %\end{table}
@@
 \begin{table}[ht]
-	\caption{Time of solving the problem using a local method}
+	\caption{Time of solving the applied problem, sec.}
 	\label{tab:12}
@@
 
-Concluding general results, one can say that the use of local methods at the final stage of search allows solving the optimization problems with higher precision.
+%\begin{table}
+	%\caption{Time of solving the problem}
+	%\label{tab:last}
+	%\center
+	%\begin{tabular}{llllllll}
+		%\hline\noalign{\smallskip}
+		%$p$ & \multicolumn{3}{l}{ Total time of solving four problems } & & \multicolumn{3}{l}{One of the temperature regimes only} \\
+		%\noalign{\smallskip} \cline{2-4} \cline{6-8} \noalign{\smallskip}
+		 %& P = 1 & P = 3 &  P = 15 & & GSA & GSA-l & GSA-pl  \\
+		%\noalign{\smallskip} \hline \noalign{\smallskip}
+		%Hooke-Jeeves & 2386.4 & 1053.1 & 713.1 & & 158.1 & 299.8  & 177.4  \\
+		%L-BFGS-B & 2571.7  & 1096.4  & 720.3 & & 158.1  & 252.5  & 165.6  \\
+%
+		%\noalign{\smallskip}\hline
+	%\end{tabular}
+%\end{table}
+
+Concluding the results, one can say that the use of local methods at the final stage of search allows solving the optimization problems with higher precision.
 In many cases, it is enough to use sequential local methods.
-However, for the problems with complex gully structure of the objective function require parallelizing the whole search process including the local stage as well.
+However, the problems with ravine-type objective function require parallelizing the whole search process including the local stage as well.
 In the present work, two local optimization algorithms with the parallelization variants for shared memory as well as for distributed one are described.

### 2022-04-15 2022/RSD/paper.tex
minor corrections

@@
 The main part of the paper has the following structure.
-In Section 2, the global optimization problem statement is presented and the properties of the objective function allowing developing the efficient global minimum search algorithms are described.
-In Section 3, the scheme of the parallel global optimization algorithm is presented and the theoretical estimates of its speedup are given.
-In Section 4, two local optimization methods (BFGS method and Hooke-Jeeves one) are considered, which can be applied for refining the solution found.
+In Section \ref{sec:problem}, the global optimization problem statement is presented and the properties of the objective function allowing developing the efficient global minimum search algorithms are described.
+In Section \ref{sec:global}, the scheme of the parallel global optimization algorithm is presented and the theoretical estimates of its speedup are given.
+In Section \ref{sec:local}, two local optimization methods (BFGS method and Hooke-Jeeves one) are considered, which can be applied for refining the solution found.
 Possible parallelization schemes for these methods are discussed.
-Section 5 contains the results of the computational experiments conducted using both the test problems and the applied ones.
-Section 6 concludes the paper.
+Section \ref{sec:results} contains the results of the computational experiments conducted using both the test problems and the applied ones.
+% Section 6 concludes the paper.
 
-\section{Global optimization problem statement}
+\section{Global optimization problem statement}\label{sec:problem}
 
@@
 
-In the theory, the Peano-Hilbert curve is a limit object.
+In theory, the Peano-Hilbert curve is a limit object.
 Therefore, one needs to construct an approximation to the theoretical curve for the practical application.
@@
 
-The use of the space-filling curves allows reducing the multidimensional problem \ref{problemN} to a one-dimensional problem and applying the efficient one-dimensional optimization methods to solve it, i.e.
+The use of the space-filling curves allows reducing the multidimensional problem (\ref{problemN}) to a one-dimensional problem and applying the efficient one-dimensional optimization methods to solve it, i.e.
 \begin{equation} \label{problem_ref}
@@
 \end{equation}
-Keeping the property of limited relative differences of the objective function is an important property of using the evolvents.
-If the function $\varphi(y)$ in the domain $D$ satisfies the Lipschitz condition, the function $\varphi(y(x))$ in the interval $[0,1]$ will satisfy the uniform H{\"o}lder condition
+Keeping the property of limited relative differences of the objective function is an important feature of the evolvents.
+If the function $\varphi(y)$ in the domain $D$ satisfies the Lipschitz condition, the function $\varphi(y(x))$ in the interval $[0,1]$ will satisfy the H{\"o}lder condition
 \begin{equation} \label{holder_ref}
@@
 \begin{equation} \label{minim_fun_ref}
-f(x) = \varphi(y(x)), x \in [0,1],
+f(x) = \varphi(y(x)), \; x \in [0,1],
 \end{equation}
@@
 
-Step 5. Carry out new trials in parallel at the points $x^{k+j}$, $1 \leq j \leq p$, computed by the formulae
+Step 5. Carry out new trials in parallel at the points $x^{k+j}$, $1 \leq j \leq p$, computed by the formula
 $$ x^{k+j} = \frac{x_{t_j} + x_{t_j - 1}}{2} - sign(z_{t_j} - z_{t_j - 1}) \frac{1}{2r} {\Bigg[\frac{\left| z_{t_j} - z_{t_j - 1} \right|}{\mu}\Bigg]}^{N}.$$
@@
 Step 6. Check the stopping criterion $\Delta_{t_j} \leq \varepsilon$, where $t_j$ are from Step 4 and $\varepsilon > 0$ is the predefined accuracy. 
-After completing the algorithm work, take as the estimate of the solution of the problem \ref{problemN} 
+After completing the algorithm work, take as the estimate of the solution of the problem (\ref{problemN})
 $$ f_k^* = \min\limits_{1 \leq i \leq k} f(x^i), \; x_k^* = \arg \min\limits_{1 \leq i \leq k} f(x^i).$$
 
-The theoretical substantiation of the presented algorithm (denoted hereafter as \textit{global search algorithm} or GSA) see in more details in \cite{Strongin2000}.
+The theoretical basis of the presented algorithm (denoted hereafter as \textit{global search algorithm} or GSA) see in more details in \cite{Strongin2000}.
 Note briefly that the characteristics of the intervals (\ref{R}) used in the algorithm can be considered as a quantitative measure how particular interval is promising from the viewpoint of finding the global minimum inside it.
@@
 \begin{equation} \label{par_trl_ref}
-s(p) = \frac{n(1)p}{n(p)}
+s(p) = \frac{n(1) \cdot p}{n(p)}
 \end{equation}
@@
 \textbf{Consequence}.
-Let the objective function $f(x)$ have $Q$ points of local minima $\{x_1^{\prime},...,x_Q^{\prime}\}$, which the condition \eqref{first_s_ref} is fulfilled for, and let there exist the trial points $y^{n_i}$, $1 \leq i \leq Q$ such as
+Let the objective function $f(x)$ have $Q$ points of local minima $\{x_1^{\prime},...,x_Q^{\prime}\}$, which the condition \eqref{first_s_ref} is fulfilled for, and let there exist the trial points $y^{n_i}$, $1 \leq i \leq Q,$ such that
 \begin{gather} 
@@
 
-The consequence from theorem plays a special role in solving the multidimensional problems reduced to the one-dimensional ones using the evolvent $y(x)$.
+The consequence from the theorem plays a special role in solving the multidimensional problems reduced to the one-dimensional ones using the evolvent $y(x)$.
 The evolvent $y(x)$, being an approximation to the Peano curve, has the effect of ``splitting'' of the global minimum point $y^* \in D$ into several (up to $2^N$) preimages in the interval [0,1].
@@
 
-Now, let us consider the case when after the stage of the search for the rough approximation to the global solution, the stage of its local refinement is performed.
+Now, let us consider the case when, after the stage of the search for the rough approximation to the global solution, the stage of its local refinement is performed.
 Assume that the fraction of the search trials necessary for the local refinement of the solution makes $\alpha$ of total number of trials.
@@
 
-The general scheme of the parallel computing for all the local methods employed is presented in Fig.
+%The general scheme of the parallel computing for all the local methods employed is presented in Fig.
 \#.
 The objective function values can be computed on a single node using shared memory (OpenMP), on a cluster with distributed memory (MPI), or on accelerator using CUDA.
+Thus, in both considered local methods the parallelization scheme will be generally the same.
+The vector of objective function values can be computed in parallel on a single node using shared memory (OpenMP), on a cluster with distributed memory (MPI), or on accelerator using CUDA.
 
-\section{Results of numerical experiments}
+\section{Results of numerical experiments}\label{sec:results}
 
@@
 \begin{gather}
-    \varphi(y) = \sum_{i=1}^{N-1} \left((1-y_i)^2 + 100(y_{i+1} - y_i^2)^2\right) \nonumber \\
+    \varphi(y) = \sum_{i=1}^{N-1} \left[(1-y_i)^2 + 100(y_{i+1} - y_i^2)^2\right], \nonumber \\
     D = \{ y \in R^N: -25 \leq y_j \leq 25, 1 \leq j \leq N \}. \nonumber
@@
 		\hline
-		N & Problem class & P = 1 & P = 2 & P = 4 & P = 8 & P = 16 \\
+		$N$ & Problem class & $p=1$ & $p=2$ & $p=4$ & $p=8$ & $p=16$ \\
 		\hline 
-		    \multirow{2}{*}{4} & hard & 36528 & 18651 & 9854 & 4808 & 2260 \\ \cline{2-7}
-		                       & simple & 9995 & 5199 & 2706 & 1333 & 630 \\
+		    \multirow{2}{*}{4} & Hard & 36528 & 18651 & 9854 & 4808 & 2260 \\ \cline{2-7}
+		                       & Simple & 9995 & 5199 & 2706 & 1333 & 630 \\
 		\hline
-		    \multirow{2}{*}{5} & hard & 266420 & 149109 & 90894 & 39873 & 22206 \\ \cline{2-7}
-		                       & simple & 29853 & 12913 & 7296 & 3618 & 2220 \\
+		    \multirow{2}{*}{5} & Hard & 266420 & 149109 & 90894 & 39873 & 22206 \\ \cline{2-7}
+		                       & Simple & 29853 & 12913 & 7296 & 3618 & 2220 \\
 		\hline

### 2022-04-15 2022/RSD/paper.tex
Обновлено описание Метода Хука-Дживса

@@
 
-if $\varphi(\bar{y} + he^i) < \varphi(\bar{y})$ then set $\bar{y} = \bar{y} + he^i$ else if $\varphi(\bar{y} - he^i) < \varphi(y)$ then $\bar{y} = y - he^i.$
+\>\>\>\> if $\varphi(\bar{y} + he^i) < \varphi(\bar{y})$ then
+
+\>\>\>\>\>\>\>\> set $\bar{y} = \bar{y} + he^i$
+
+\>\>\>\> else if $\varphi(\bar{y} - he^i) < \varphi(y)$ then
+
+\>\>\>\>\>\>\>\> set $\bar{y} = y - he^i.$
 
@@
 Step 0.
-Set the starting point $t^0$, the parameter of the method $\varepsilon > 0$, the parameter of constructing the configuration $h\gg\varepsilon$, and the parameter $\alpha=2$ of the increasing of the step, e.g.
-$\alpha=2$.
+Set the starting point $t^0$, the parameter of the method $\varepsilon > 0$, the parameter of constructing the configuration $h\gg\varepsilon$, and the parameter $\alpha = 2$ of the increasing of the step, e.g.
+$\alpha = 2$.
 
@@
 
-Step 3.
-If $\varphi(t^{k+1}) < \varphi(t^{k})$ then $k:=k+1$ and go to Step 4, else if $h \leq \varepsilon$, stop the search, if $h > \varepsilon$ then further actions depend on the  way how the point $y^{k+1}$ was constructed: whether the configuration was constructed using the step by sample (in this case, $k > 0$) or it was constructed from the point $t^0$(in this case, $k=0$).
-If $k=0$, cut $h$ in half ($h=h/2$) and go to Step 1.
-If $k>0$, set $t^0 = t^k, k=0$ and also go to Step 1.
+Step 3. Determine the next step
 
-Step 4. Perform the step by direction $y^{k+1} = t^k + \alpha (t^k - t^{k-1})$ and go to Step 2.
+\>\>\>\> if $\varphi(t^{k + 1}) < \varphi(t^{k})$ then
+
+\>\>\>\>\>\>\>\> $k = k + 1$ and go to Step 4,
+
+\>\>\>\> else
+
+\>\>\>\>\>\>\>\> if $h \leq \varepsilon$ then
+
+\>\>\>\>\>\>\>\>\>\>\>\>  stop the search
+
+\>\>\>\>\>\>\>\> else if $h > \varepsilon$ then
+
+\>\>\>\>\>\>\>\>\>\>\>\> if $k = 0$ then
+
+\>\>\>\>\>\>\>\>\>\>\>\>\>\>\>\> $h = h / 2$ and go to Step 1.
+
+\>\>\>\>\>\>\>\>\>\>\>\> else if $k > 0$ then
+
+\>\>\>\>\>\>\>\>\>\>\>\>\>\>\>\> set $t^0 = t^k, k = 0$ and go to Step 1.
+
+Step 4. Perform the step by direction $y^{k + 1} = t^k + \alpha (t^k - t^{k - 1})$ and go to Step 2.
 

### 2022-04-15 2022/RSD/paper.tex
before submission

@@
 %
-\title{Analysis and elimination of bottlenecks in parallel algorithm for solving global optimization problems}
+\title{Analysis and elimination of bottlenecks\\ in parallel algorithm for solving\\ global optimization problems}
 %
@@
 It is constructed so that the direction $(\bar{y} - y)$ is that of the decreasing of the function $\varphi$ in the nearness of $y$.
-To describe the operator $F$, let us introduce the following notations: $e^i$ is the $i^{th}$ unit coordinate vector and $h$ is the parameter determining the magnitude of the coordinate translation.
+To describe the operator $F$, let us introduce the following notations: $e^i$ is the $i^{th}$ unit coordinate vector and $h$ is the parameter determining the magnitude of the coordinate movement.
 Then, the transition from $y$ to $\bar{y}$ (i.e.
@@
 Step 0.
-Set the starting point $t^0$, the parameter of the method $\varepsilon > 0$, the parameter of constructing the configuration $h\gg\varepsilon$, and the parameter $\alpha = 2$ of the increasing of the step, e.g.
+Set the starting point $t^0$, the parameter of the method $\varepsilon > 0$, the parameter of constructing the configuration $h\gg\varepsilon$, and the parameter $\alpha$ of the increasing of the step, e.g.
 $\alpha = 2$.
@@
 
-\>\>\>\>\>\>\>\> if $h \leq \varepsilon$ then
-
-\>\>\>\>\>\>\>\>\>\>\>\>  stop the search
+\>\>\>\>\>\>\>\> if $h \leq \varepsilon$ then stop the search
 
@@
 
-Step 4. Perform the step by direction $y^{k + 1} = t^k + \alpha (t^k - t^{k - 1})$ and go to Step 2.
+Step 4. Move by direction $y^{k + 1} = t^k + \alpha (t^k - t^{k - 1})$ and go to Step 2.
 
 The meaning of the actions performed within the steps 2, 3, and 4 can be explained as follows. 
-Step 4 was introduced in order to enable the method to increase the shift magnitude within a single iteration rapidly in the case when the point $t^k$ is located far away from the solution.
+Step 4 was introduced in order to enable the method to rapidly increase the shift magnitude within a single iteration in the case when the point $t^k$ is located far away from the solution.
 At that, prior to make $y^{k+1}$ current point of the iteration, a configuration is constructed from the point $y^{k+1}$.
-By these means, in the case of obtaining a point with $\varphi(t^{k+1})< \varphi(t^k)$, the next step by sample, in general, will be performed in a changed direction (relative to previous one) that allows adapting the method to variations in the function relief.
+In the case of obtaining a point with $\varphi(t^{k+1})< \varphi(t^k)$, the next step, in general, will be performed in a changed direction (relative to previous one) that allows adapting the method to variations in the function relief.
 
 Finally, upon obtaining a value $\varphi(t^{k+1}) \geq \varphi(t^k)$ at Step 3, an attempt to run the algorithm again from the point $t^0=t^k$ (better than the one found) is undertaken.
-At that, if such an attempt hasn’t been undertaken yet, the parameter $h$ is not changed, but if it has been undertaken already, $h$ is cut in half first.
+At that, if such an attempt hasn't been undertaken yet, the parameter $h$ is not changed, but if it has been undertaken already, $h$ is cut in half first.
 
@@
 
-Fig.~\ref{fig:RosenbrocNT} presents the averaged numbers of iterations, which Hooke-Jeeves and BFGS methods executed for minimization of Rosenbrock function with the dimensionalities from 2 to 30.
+Fig.~\ref{fig:RosenbrocNT} presents the averaged numbers of iterations, which Hooke-Jeeves and BFGS methods executed for minimization of Rosenbrock function with the dimensionalities from 5 to 30.
 Fig.~\ref{fig:RosenbrocSpeedup} presents corresponding speedups in iterations.

### 2022-04-15 2022/Sensors/kNN.tex

@@
 As a result, when using Peano curve type mappings, close objects $y^\prime$, $y^{\prime\prime}$ in the $N$-dimensional space can have their quite distant corresponding preimages $x^\prime$, $x^{\prime\prime}$ on the interval $[0,1]$.
-In Fig.
-1 the green dots show two objects that are close in the two-dimensional space, while their preimages on the one-dimensional scale are far away from each other.
+In Figure~\ref{fig:Peano} the green dots show two objects that are close in the two-dimensional space, while their preimages on the one-dimensional scale are far away from each other.
 The blue dots correspond to two objects, the distance between which is retained when changing to the one-dimensional scale.
@@
 \includegraphics[width=0.5\linewidth]{fig1.png}
-\caption{Peano curve}
+\caption{Peano curve}\label{fig:Peano}
 \label{fig1}
@@
 \end{equation} 
-map the interval $[0,1]$ onto the corresponding hypercubes  $D_l, \; 1 \leq l \leq L$ (the broken line in Fig.~\ref{fig2} shows the image of the interval $[0,1]$, which is obtained by using the evolvent $y^0(x), x \in [0,1]$).
+map the interval $[0,1]$ onto the corresponding hypercubes  $D_l, \; 1 \leq l \leq L$ (the broken line in Figure~\ref{fig2} shows the image of the interval $[0,1]$, which is obtained by using the evolvent $y^0(x), x \in [0,1]$).
 
@@
 
-Since the hypercube $D$ from (\ref{D}) belongs to the common part of the family of hypercubes (\ref{hypercubes}) (the boundary of the hypercube $D$ is highlighted in Fig.~\ref{fig2}), then, by introducing an additional function
+Since the hypercube $D$ from (\ref{D}) belongs to the common part of the family of hypercubes (\ref{hypercubes}) (the boundary of the hypercube $D$ is highlighted in Figure~\ref{fig2}), then, by introducing an additional function
 \[
@@
 
-As an illustration, Fig.
-\ref{fig2} shows a family of space-filling curves in extended domains $D_l$, where a square with a dark border highlights the hypercube $D$ belonging to each extended domain.
+As an illustration, Figure~\ref{fig2} shows a family of space-filling curves in extended domains $D_l$, where a square with a dark border highlights the hypercube $D$ belonging to each extended domain.
 The dots indicate proximate objects in the domain $D$, the distance between which will vary depending on the one-dimensional scale used.
@@
 
-The running times used by the different methods (kNN-ME, procedure A, kNN-ME, procedure B, kNN-KD) for DS-Random-1 and DS-Random-2 datasets with $N = 2$ are shown in Fig.
-\ref{fig3}.
+The running times used by the different methods (kNN-ME, procedure A, kNN-ME, procedure B, kNN-KD) for DS-Random-1 and DS-Random-2 datasets with $N = 2$ are shown in Figure~\ref{fig3}.
 The abscissa axis represents the number of nearest neighbours that the algorithm is searching for, the ordinate axis represents the running time of the method.
@@
 
-Running times of the methods for nearest neighbours search with $N = 3, 4$ are given in Figs. \ref{fig4}, \ref{fig5} respectively.
+Running times of the methods for nearest neighbours search with $N = 3, 4$ are given in Figures \ref{fig4}, \ref{fig5} respectively.
 
@@
 
-The time required to find $K$ nearest neighbours on the CarEvoluation dataset is shown in Fig.~\ref{fig6a}. In this experiment, the number of evolvents used simultaneously was 7. 
+The time required to find $K$ nearest neighbours on the CarEvoluation dataset is shown in Figure~\ref{fig6a}. In this experiment, the number of evolvents used simultaneously was 7. 
 \begin{figure}

### 2022-04-15 2022/RSD/paper.tex
after submission

@@
 Two stages of solving the problem were distinguished: the search of a rough approximation to the global solution and its local refining.
-The novelty of the parallel algorithm considered in the present pa-per consists in the parallelization of the whole computational process -- both at the global search stage and at the one of the local refining of the solution found.
+The novelty of the parallel algorithm considered in the present paper consists in the parallelization of the whole computational process, both at the global search stage and at the one of the local refining of the solution found.
 The theoretical estimates of the algorithm parallelization efficiency are presented.
@@
 
-\keywords{First keyword  \and Second keyword \and Another keyword.}
+\keywords{Multiextremal functions \and Lipschitz condition \and Global optimization \and Local optimization \and Parallel algorithms}
 \end{abstract}
@@
 
-The main requirement to forming the corrections $U^k$ of particular form is the secant equation $B^{k+1} (y^{k+1} - y^k )= \nabla \varphi(y^{k+1})- \nabla \varphi(y^k)$, which is performed for all quasi-Newton methods.
+The main requirement to forming the corrections $U^k$ of particular form is the condition
+$$B^{k+1} (y^{k+1} - y^k )= \nabla \varphi(y^{k+1})- \nabla \varphi(y^k),$$
+that holds true for all quasi-Newton methods.
 For more precise determination of the matrix $B^{k+1}$, it is necessary to impose additional requirements on the Hessian approximation.

### 2022-04-29 2022/LION/main.tex
Lion добавлен график изменения HV при разном числе итераций
Обновил график
include three new rerefences

@@
 One of such problem classes to which ML can be applied is the class of multicriterial optimization (MCO) models.
-In this case, ML is used, as a rule, in combination with metaheuristic algorithms \cite{Subraveti2019} that in the case of multiextremal criteria lose in efficiency to deterministic methods \cite{Kvasov2018,Sergeyev2018}.
+In this case, ML is used, as a rule, in combination with metaheuristic algorithms \cite{Ma2014,Subraveti2019,Talbi2015} that in the case of multiextremal criteria lose in efficiency to deterministic methods \cite{Kvasov2018,Sergeyev2018}.
 
@@
 
-Computational testing was carried out on the``Lobachevsky'' supercomputer of Nizhny Novgorod State University in the environment of the Globalizer software system \cite{globalizerSystem}.
+Computational testing was carried out on the ``Lobachevsky'' supercomputer of Nizhny Novgorod State University in the environment of the Globalizer software system \cite{globalizerSystem}.
 
@@
 Other algorithm's parameters were the same.
-The averaged results can be found in Table \ref{tab:2}.
+The averaged results can be found in Table~\ref{tab:2}.
+\textcolor[rgb]{1,0,0}{Fig.~\ref{fig:2} additionally shows the average value of HV index change depending of the number of trials.}
 
@@
 
-As shown in the table, embedding the machine learning into the global search algorithm results in a significant reduction of the number of trials (almost 3 times) with maintaining close values of indicators HV and DU.
+\begin{figure}[ht]
+\center
+\includegraphics[width=0.7\textwidth]{fig2.png}
+\caption{Efficiency of the multicriterial algorithms compared} \label{fig:2}
+\end{figure}
+
+\textcolor[rgb]{1,0,0}{The results show that embedding the machine learning into the global search algorithm leads to }a significant reduction of the number of trials (almost 3 times) with maintaining close values of indicators HV and DU.
 

### 2022-04-30 2022/PAVT/PAVT2022.tex
final
after proofreading

@@
 %\title{Solving the Inverse Problems of Chemical Kinetics Using the Asynchronous Global Optimization Algorithm}
-\title{Kinetic Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as Catalyst Using the Asynchronous Global Optimization Algorithm}
+\title{Kinetic Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as a Catalyst Using the Asynchronous Global Optimization Algorithm}
 \titlerunning{Kinetic Modeling of Isobutane Alkylation}  % abbreviated title (for running head)
@@
 
-\institute{$^1$Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia\\$^2$Ufa State Petroleum Technological University, Ufa, Russia \\$^3$Novosibirsk State University, Novosibirsk, Russia\\$^4$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
+\institute{$^1$Institute of Petrochemistry and Catalysis -- Subdivision of Ufa Federal Research Centre of the RAS, Ufa, Russia\\$^2$Ufa State Petroleum Technological University, Ufa, Russia \\$^3$Novosibirsk State University, Novosibirsk, Russia\\$^4$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
 \email{leniza.enikeeva@yandex.ru},
@@
 
-The paper considers the application of parallel computing technology to the simulation of catalytic chemical reaction, which is widely used in the modern automobile industry to produce gasoline with high octane number.
-As a chemical reaction, the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid is assumed.
-To simulate a chemical process, it is necessary to develop a kinetic model of the process, that is, to determine the kinetic parameters.
-To do this, the inverse problem of chemical kinetics is solved, which predicts the values of kinetic parameters based on laboratory data.
+The paper considers the application of parallel computing technology to the simulation of a catalytic chemical reaction, which is widely used in the modern automobile industry to produce gasoline with a high octane number.
+As a chemical reaction, the process of alkylation of isobutane with mixed C4 olefins, catalyzed by sulfuric acid, is assumed.
+To simulate a chemical process, it is necessary to develop a kinetic model of the process, i.e., to determine the kinetic parameters.
+To do this, the inverse problem of chemical kinetics is solved; it predicts the values of the kinetic parameters based on laboratory data.
 From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem.
 A parallel asynchronous information-statistical global search algorithm was used to solve it.
-The use of the asynchronous algorithm has significantly reduced the search time to find the optimum.
+The use of the asynchronous algorithm significantly reduced the search time to find the optimum.
 The found optimal parameters of the model made it possible to adequately simulate the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid.
@@
 %обновить ключевые слова
-\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Chemical kinetics $\cdot$ Inverse problems }
+\keywords{Global optimization $\cdot$ Multi-extremal functions $\cdot$ Parallel computing $\cdot$ Chemical kinetics $\cdot$ Inverse problems }
 \end{abstract}
@@
 Sulfuric acid alkylation of isobutane with olefins makes it possible to obtain a high-octane component of gasoline with a minimum content of aromatic hydrocarbons.
-The alkylate, which is produced by the alkylation of isobutane with C3 -- C5 olefins in the presence of strong acid, has the advantages of high octane number, low vapor pressure, and zero content of olefins and aromatics that allow it to be a desirable blending component for high-quality gasoline.
+The alkylate, which is produced by alkylation of isobutane with C3 -- C5 olefins in the presence of strong acid, has the advantages of a high octane number, low vapor pressure, and zero content of olefins and aromatics, making it a desirable blending component for high-quality gasoline.
 Alkylates will continue to act as a desirable blending component for high-quality gasoline as the quality of gasoline continues to increase \cite{cao2019}.
-Therefore, it is a significant process of the modern refinery.
-To optimize the chemical process in industry, it is necessary to first develop its model, which in this case means building a mathematical model of the chemical process and then its kinetic model, that is, numerically calculate the kinetic constants of the reaction.
+Therefore, it is a significant process for a modern refinery.
+To optimize the chemical process in industry, it is necessary to develop first its model, which in this case means building a mathematical model of the chemical process, and then its kinetic model, i.e., to numerically calculate the kinetic constants of the reaction.
 
 %Стыковочный абзац
-As a rule, it is impossible to find out the kinetic constants of the reactions analytically.
+As a rule, it is impossible to find out the kinetic constants of reactions analytically.
 Therefore, there is a need in the development and application of numerical methods for finding the kinetic constants (see, e.g., \cite{Zaynullin2020,Enikeeva2020,Koledina2019,Enikeev2020,uskovfibrous,enikeeva2021,en13133393}).
-In this case, the quality criteria of the solution found (objective function) haven't an explicitt analytical description but allows an algorithmic representation and requires considerable computaton resources.
-Moreover, in the inverse problems of chemical kinetics the objective function can be multiextremal essentially, i.e.
-can have many local extrema along with the global one.
+In this case, the quality criteria of the solution found (objective function) do not have an explicit analytical description, but enable an algorithmic representation and require considerable computational resources.
+Moreover, in the inverse problems of chemical kinetics, the objective function can be essentially multi-extremal, i.e., can have many local extrema along with the global one.
 
 %Часть ННГУ
-The numerical methods for solving such multiextremal problems (global optimization methods) differ significantly from the local search ones (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}).
-As a rule, the local optimization methods cannot escape a local extremum attractione region and don't find the global optimum.
-At the same time, the use of the model parameters corresponding to the local solution found may appear to be insufficient since the global solution may provide a considerable advantage as compared to the local ones.
+Numerical methods for solving such multi-extremal problems (global optimization methods) differ significantly from local search methods (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}).
+As a rule, local optimization methods cannot escape the local extremum attraction region and do not find the global optimum.
+At the same time, the use of model parameters corresponding to the found local solution may appear to be insufficient since the global solution can provide a considerable advantage over local ones.
 
-The diversity of the global optimization problems arising entails various approaches to solving these ones.
-The methods of solving the global optimization problems can be divided into two classes: the metaheuristic methods and the deterministic ones.
-The metaheuristic algorithms, as a rule, are based on the simulation of the processes going in the nature.
-Some examples of the metaheuristic algorithms are simulated annealing, evolution and genetic algorithms, etc.
-(see e.g.
-\cite{Battiti2009,Eiben2015}).
-Because of relative simplicity, the metaheuristic algorithms are more popular among the researchers than the deterministic methods.
-However, a problem solution found by a metaheuristic algorithm is, generally speaking, a local one and may be far from the global solution \cite{Kvasov2018}.
+The diversity of emerging global optimization problems entails various approaches to their solution.
+Methods for solving global optimization problems can be divided into two classes: metaheuristic and deterministic.
+Metaheuristic algorithms are usually based on the simulation of processes occurring in nature.
+Some examples of metaheuristic algorithms are simulated annealing, evolution and genetic algorithms, etc.
+(see, e.g., \cite{Battiti2009,Eiben2015}).
+Due to their relative simplicity, metaheuristic algorithms are more popular among researchers than deterministic methods.
+However, the problem solution found by the metaheuristic algorithm is, generally speaking, local and may be far from the global solution \cite{Kvasov2018}.
 
-The possibility to construct deterministic global search methods different from the grid search and from the metaheuristic methods is related to the availability of and taking into account some {\it a priori} assumption on the properties of the problem functions.
-Such assumptions play a key role in the development of efficient global optimization algorithms and serve as main mathematical tools for estimating the global solutions.
+The possibility to construct deterministic global search methods different from grid search and metaheuristic methods is related to the availability and consideration of some {\it a priori} assumption on the properties of problem functions.
+Such assumptions play a key role in the development of efficient global optimization algorithms and serve as the main mathematical tool for estimating global solutions.
 
-An assumption on limited relative variations of the objective function values is one of natural assumptions of a problem.
-Such an assumption is related to the ratio of the function increment to respective increment of its argument, which is usually limited by some threshold defined by a limited energy of variations in the simulated system.
-In this case the functions are known as Lipschitz ones and the problem itself is called the Lipschitz global optimization problem.
+The assumption on limited relative variations of objective function values is one of the natural assumptions of the problem.
+Such an assumption is related to the ratio of the function increment to the respective increment of its argument, which is usually limited by some threshold defined by the limited energy of variations in the simulated system.
+In this case, the functions are known as Lipschitz ones, and the problem itself is called the Lipschitz global optimization problem.
 
-This paper presents the results of application of parallel Lipschitz optimization methods for solving the inverse problems of chemical kinetics.
+This paper presents the results of applying parallel Lipschitz optimization methods for solving the inverse problems of chemical kinetics.
 The main part of the paper has the following structure.
 The description of the mathematical model of the investigated chemical reaction is presented in Section \ref{Sec_math_mod}.
-The formal statement of the Lipschitz global optimization problems and the asynchronous parallel algorithm for solving these ones are described in Section \ref{Sec_GSA}.
-The results of the numerical solving the inverse problem of chemical kinetics are discussed in Section \ref{Sec_Exp}.
+The formal statement of Lipschitz global optimization problems and the asynchronous parallel algorithm for solving them are described in Section \ref{Sec_GSA}.
+The results of the numerical solution of the inverse problem of chemical kinetics are discussed in Section \ref{Sec_Exp}.
 
@@
 The initial conditions are $t = 0, c_1 = c_1^0; c_2=c_2^0; c_3 = 0; c_4 = 0; c_5= 0; c_6 = 0; c_7 = 0; c_8 = 0; c_9 = 0; c_{10} = 0; c_{11}=c_{11}^0; c_{12} = c_{12}^0$.
-The corresponding species in eqs (\ref{eq:one})--(\ref{eq:twelve}) are 1, iC4H8; 2, iC4, 3, iC4+; 4, TMPs+; 5, DMHs+; 6, TMPs; 7, DMHs; 8, HEs; 9, iCx+; 10, iCy=; 11, 2-C4H8; 12, 1-C4H8.
+The corresponding species in equations (\ref{eq:one})--(\ref{eq:twelve}) are 1, iC4H8; 2, iC4, 3, iC4+; 4, TMPs+; 5, DMHs+; 6, TMPs; 7, DMHs; 8, HEs; 9, iCx+; 10, iCy=; 11, 2-C4H8; 12, 1-C4H8.
 
-As experimental data, information is taken from the article \cite{cao2019}, which represent a change in the concentrations of reaction components over time at different temperatures; as an example of experimental data at a temperature of 276.2 K, the data in Table \ref{table1} are presented.
+Information that represents a change in the concentrations of the reaction components over time at different temperatures is taken as experimental data from \cite{cao2019}; the data in Table \ref{table1} are presented as an example of experimental data at a temperature of 276.2 K.
 
 \begin{table}
-\caption{Experimental Data}
+\caption{Experimental data}
 \label{table1}
@@
 
-However, it is necessary to take into account the fact that the reaction rate constants $k_1, k_2, ..., k_{15}$ included in the equations (\ref{eq:one})--(\ref{eq:twelve}) are parameters depending on the reaction temperature, and this dependence is the Arrhenius equation and has the following form:
+However, it is necessary to take into account the fact that the reaction rate constants $k_1, k_2, ..., k_{15}$ included in equations (\ref{eq:one})--(\ref{eq:twelve}) are parameters depending on the reaction temperature, this dependence is the Arrhenius equation and has the following form:
 
@@
 \end{equation}
-where $k_i(T)$ is a constant of $i$-th stage of reation rate, $k_i^0$ is a pre-exponential factor of $i$-th reation stage, $E_i$ is an activation energy, J/mol, $R$ is the universal gas constant, J $\cdot$ (K$\cdot$mol), $T$ is the temperature, K.
-Thus, in order to fully develop the kinetic model of the reaction, it is necessary to calculate the activation energies $E_i$ and pre-exponential factors $k_i^0$ of all stages of the chemical reaction.
+where $k_i(T)$ is the constant of the $i$-th stage of the reaction rate, $k_i^0$ is the pre-exponential factor of the $i$-th reaction stage, $E_i$ is the activation energy, J/mol, $R$ is the universal gas constant, J $\cdot$ (K$\cdot$mol), $T$ is the temperature, K.
+Thus, to fully develop the kinetic model of the reaction, it is necessary to calculate the activation energies $E_i$ and the pre-exponential factors $k_i^0$ of all stages of the chemical reaction.
 There are two formulations of the problems of searching for kinetic parameter data $E$ and $k^0$.
-The first one is to solve the inverse problem of selecting the kinetic parameters included in the equation (\ref{eq:arren}), which during the solution allow us to calculate all the reaction rate constants $k_i$.
-Then, with the found rate constants of the reaction stages, the system of differential equations (\ref{eq:one})--(\ref{eq:twelve}) is solved, then the calculated concentrations are compared with the corresponding experimental data.
+The first one is to solve the inverse problem of selecting the kinetic parameters, included in equation (\ref{eq:arren}), which during the solution allow us to calculate all the reaction rate constants $k_i$.
+With the found rate constants of the reaction stages, the system of differential equations (\ref{eq:one})--(\ref{eq:twelve}) is solved, then the calculated concentrations are compared with the corresponding experimental data.
 Mathematically, this problem has the following formulation: it is necessary to minimize the following objective function
@@
 \end{equation}
-where $c_{ijk}^{exp}$ and $c_{ijk}^{calc}$ are the experimental and calculated values of $k$-th observed component at $i$-th experiment, respectively, $I$ is the the number of observed temperatures for the reaction, ($I = 4$ in this case), $K$ is the number of experiments conducted at one temperature ($K = 3$), $M$ is the number of observed components of the reaction ($M = 2$).
+where $c_{ijk}^{exp}$ and $c_{ijk}^{calc}$ are the experimental and calculated values of the $k$-th observed component at the $i$-th experiment, respectively, $I$ is the number of observed temperatures for the reaction, ($I = 4$ in this case), $K$ is the number of experiments conducted at one temperature ($K = 3$), $M$ is the number of observed components of the reaction ($M = 2$).
 
-The second formulation of the inverse problem of chemical kinetics implies the search for the rate constants of the stages $k_i$ included in the system (\ref{eq:one})--(\ref{eq:twelve}), separately for each temperature, then on the basis of the Arrhenius equation~(\ref{eq:arren}), the activation energies $E_i$ and the pre-exponential multipliers $k_i^0$ are calculated using the least squares method.
-Mathematically, this statement of the problem coincides with the equation (\ref{eq:objective_func1}), except for the first summation by temperatures:
+The second formulation of the inverse problem of chemical kinetics implies the search for the rate constants of the stages $k_i$ included in the system (\ref{eq:one})--(\ref{eq:twelve}), separately for each temperature, then on the basis of Arrhenius equation~(\ref{eq:arren}), the activation energies $E_i$ and the pre-exponential multipliers $k_i^0$ are calculated using the least squares method.
+Mathematically, this statement of the problem coincides with equation (\ref{eq:objective_func1}), except for the first summation by temperatures:
 
@@
 
-Thus, both formulations of the problem (\ref{eq:objective_func1}) and (\ref{eq:objective_func2}) are optimization problems and the next section will describe the method that was used to solve these minimization problems.
+Thus, both formulations of the problem (\ref{eq:objective_func1}) and (\ref{eq:objective_func2}) are optimization problems, and the next section will describe the method used to solve these minimization problems.
 
@@
 
-As it has been already mentioned above, from the formal point of view we consider the inverse problem of chemical kinetics as a global optimization problem. 
-In the specific problem under consideration, the values of the objective function are calculated by solving stiff ODE system (\ref{eq:one})--(\ref{eq:twelve}).
-Since the right parts of the system are continuous functions with bounded derivatives, then theoretically its solution will also be continuous and bounded.
-Therefore, the discrepancy (\ref{eq:objective_func2}) will satisfy the Lipschitz condition with a priori unknown constant.
+As mentioned above, from a formal point of view, we consider the inverse problem of chemical kinetics as a global optimization problem. 
+In the specific problem under consideration, the values of the objective function are calculated by solving the stiff ODE system (\ref{eq:one})--(\ref{eq:twelve}).
+Since the right parts of the system are continuous functions with bounded derivatives, theoretically its solution will also be continuous and bounded.
+Therefore, discrepancy (\ref{eq:objective_func2}) will satisfy the Lipschitz condition with a priori unknown constant.
 
-In the general form, a problem of the class specified above can be formulated mathematically as 
+In the general form, the problem of the class specified above can be formulated mathematically as 
 follows:
@@
 \end{gather}
-where $a,b$ are given vectors, $a,b\in R^N$, and the objective function $\varphi(y)$ satisfies the Lipschitz condition
+where $a,b$ are the given vectors, $a,b\in R^N$, and the objective function $\varphi(y)$ satisfies the Lipschitz condition
 \begin{equation}\label{Lip}
@@
 
-The function $\varphi(y)$ is assumed to be multiextremal and defined in the form of ``black box'' (i.e.
-in the form of some computing procedure, the input of which the  vector of parameters is supplied into, and the corresponding function value is taken from the output).
-Moreover, each \textit{trial} (i.e.
-the computing of the function value at a point of the search domain) is assumed to be time-consuming operation.
-As it has been noted in Introduction, such problem statement corresponds to the inverse problem of chemical kinetics completely.
+The function $\varphi(y)$ is assumed to be multi-extremal and defined in the form of ``black box'' (i.e., in the form of some computing procedure, into the input of which the vector of parameters is supplied, and the corresponding function value is taken from the output).
+Moreover, each \textit{trial} (i.e., the computation of the function value at a point of the search domain) is assumed to be a time-consuming operation.
+As noted in Introduction, such a problem statement corresponds to the inverse problem of chemical kinetics completely.
 
-The Lipschitz condition (\ref{Lip}) can be utilized to estimate the global minimum of a function within an interval, and knowing the Lipschitz constant allows constructing the global search algorithms and proving the convergence conditions for these ones (see, for example, \cite{Strongin2000}).
+Lipschitz condition (\ref{Lip}) can be utilized to estimate the global minimum of a function within an interval, and knowing the Lipschitz constant allows constructing global search algorithms and proving the convergence conditions for them (see, e.g., \cite{Strongin2000}).
 
-The growth of the computation costs with increasing problem dimensionality is one of the main difficulties in solving the multidimensional global optimization problems.
-The decreasing of the number of trials at preserving the solution accuracy is possible by a complete utilization of some {\it a priori} assumptions on the objective function that leads to adaptive sequential optimization algorithms.
+The growth of computational costs with increasing the problem dimensionality is one of the main difficulties in solving multidimensional global optimization problems.
+Decreasing the number of trials at preserving the solution accuracy is possible by the complete utilization of some {\it a priori} assumptions on the objective function, which leads to adaptive serial optimization algorithms.
 
-For example, non-uniform space covering method \cite{Evtushenko2013} and simplicial partitions method \cite{Zilinskas2010} are such methods.
-These approaches were applied successfully for the development of the parallel optimization methods as well \cite{Evtushenko2009,Paulavicius2011}.
-Another adaptive approach to solving a multidimensional problem (\ref{problemN}) is its reduction to a single one-dimensional problem or to several ones followed by application of the one-dimensional algorithms.
-Such a reduction can be made, for example, using nested optimization scheme \cite{Grishagin2018} or Peano-Hilbert curves \cite{Barkalov2018}. 
+For example, the non-uniform space covering method \cite{Evtushenko2013} and the simplicial partitions method \cite{Zilinskas2010} are such methods.
+These approaches were successfully applied for the development of parallel optimization methods as well \cite{Evtushenko2009,Paulavicius2011}.
+Another adaptive approach to solving multidimensional problem (\ref{problemN}) is its reduction to a single one-dimensional problem or to several ones followed by the application of one-dimensional algorithms.
+Such a reduction can be made, for example, using the nested optimization scheme \cite{Grishagin2018} or Peano-Hilbert curves \cite{Barkalov2018}. 
 The latter approach was used in the present work.
 
-Using the continuous unambiguous mapping (Peano-Hilbert curve) $y(x)$ of the interval $[0,1]$ of the real axis on the hypercube $D$ from (\ref{D}), one can reduce a multidimensional problem (\ref{problemN}) to a one-dimensional problem
+Using the continuous unambiguous mapping (Peano-Hilbert curve) $y(x)$ of the interval $[0,1]$ of the real axis on the hypercube $D$ from (\ref{D}), one can reduce multidimensional problem (\ref{problemN}) to a one-dimensional problem
 \[
@@
 \]
-where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
+where the function $\varphi(y(x))$ will satisfy the uniform H{\"o}lder condition
 \[
@@
 
-So far, a search trial at some point $x'\in[0,1]$ will include first the construction of the image $y'=y(x')$ and then the computing the value of the function $z'=\varphi(y')$.
+So far, a search trial at some point $x'\in[0,1]$ will include first the construction of the image $y'=y(x')$ and then the computation of the value of the function $z'=\varphi(y')$.
 
@@
 In the approach proposed, the parallelization scheme corresponds to the ``master/worker'' principle.
-In the master process the global search algorithm is executed, which accumulates the search information, evaluates the Lipschitz constant for the objective function on its base, determines the new trial points and distributes these ones among the worker processes.
+In the master process, the global search algorithm is executed, it accumulates search information, evaluates the Lipschitz constant for the objective function on its base, determines new trial points and distributes them among worker processes.
 
-The worker processes receive the trial points from the master process, perform the new trials at these points, and send the trial results to the master process.
+Worker processes receive the trial points from the master process, perform new trials at these points and send the trial results to the master process.
 
-Let us assume the master process to compute one point of the next trial at every iteration and to send it to a worker process for executing the trial. 
-At the same time, the execution of the trial by the worker process is much more computational-costly operation than the choice of a new trial point by the master that excludes idle worker processes. 
-In this case (unlike the synchronous parallel algorithms), the total number of trials executed by each worker process will depend on the computation costs of executing a particular trial and cannot be estimated in advance.
+Let us assume that the master process computes one point of the next trial at each iteration and sends it to the worker process for executing the trial. 
+At the same time, the execution of the trial by the worker process is a much more computationally expensive operation than the choice of a new trial point by the master that excludes idle worker processes.
+In this case (unlike synchronous parallel algorithms), the total number of trials executed by each worker process will depend on the computational costs of executing a particular trial and cannot be estimated in advance.
 
-In the description of the parallel algorithm, let us assume $p+1$ computing processes to be at our disposal: one master process and $p$ worker ones.
+In the description of the parallel algorithm, let us assume that $p+1$ computational processes are at our disposal: one master process and $p$ worker ones.
 
 At the beginning of the search, the master process (let us assume it to be Process No 0) initiates the parallel execution of $p$ trials at $p$ different points of the search domain. 
-Two of these points are the boundary ones while the rest are the internal ones, i.e. at the points $\{y(x^1), y(x^2), ...,y(x^p)\}$ where 
+Two of these points are boundary, while the rest are internal, i.e., at the points $\{y(x^1), y(x^2), ...,y(x^p)\}$ where 
 $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.
 
-Now let us assume $k\geq 0$ trials (in particular, $k$ may be equal to 0) to be completed, and the worker processes are performing the trials at the points $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.
+Now let us assume that $k\geq 0$ trials (in particular, $k$ can be equal to 0) are completed, and worker processes perform trials at the points $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.
 
-Each worker process having completed its trial at some point (without any loss of generality, let us assume this point to be $y(x^{k+1})$ corresponding to process No 1), sends the trial result to the master process.
+Each worker process, having completed its trial at some point (without any loss of generality, let us assume this point to be $y(x^{k+1})$ corresponding to Process No 1), sends the trial result to the master process.
 In turn, the master process selects a new trial point $x^{k+p+1}$ for the worker process according to the rules described below.
@@
 \]
-which the trials have been already started at but haven't been completed yet.
+at which the trials have already started, but have not yet been completed.
 
-Step 1. Renumber set of preimages of the trial points 
+Step 1. Renumber the set of preimages of the trial points 
 \[
@@
 \]
-containing all the preimages, which the trials either have been completed or are being performed at in the increasing order (by the lower index), so that
+containing all preimages at which the trials either have been completed or underway in the increasing order (by the lower index) so that
 \[
@@
 where $z_i=\varphi(y(x_i))$ if $x_i \notin I_k, \; 1\leq i \leq k+p$.
-The values $z_i$ at the points $x_i \in I_k$ are undefined since the trials at the points $x_i \in I_k$ haven't been completed yet.
+The values $z_i$ at the points $x_i \in I_k$ are undefined since the trials at the points $x_i \in I_k$ have not yet been completed.
 If the value of $M$ equals 0, then set $M=1$.
@@
 Step 3.
-Juxtapose each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$ to a quantity $R(i)$, which is called a characteristic of the interval and is computed according to the formula
+Juxtapose each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$ to the quantity $R(i)$, which is called the characteristic of the interval and is computed according to the formula
 \begin{equation}\label{R}
@@
 
-Step 4. Select the interval $[x_{t-1},x_t]$, which the maximum characteristic corresponds to, i.e.
+Step 4. Select the interval $[x_{t-1},x_t]$, which the maximum characteristic corresponds to, i.e.,
 \[
@@
 
-Step 5. Define the new trial point $y^{k+p+1}=y(x^{k+p+1})$, the preimage of which is $x^{k+p+1} \in (x_{t-1},x_t)$ according to the formula
+Step 5. Define a new trial point $y^{k+p+1}=y(x^{k+p+1})$, the preimage of which is $x^{k+p+1} \in (x_{t-1},x_t)$ according to the formula
 \[
@@
 
-Upon computing the next trial point, the master process adds it to the set $I_k$ and sends it to the worker process, which initiates the new trial at this point.
+Upon computing the next trial point, the master process adds it to the set $I_k$ and sends it to the worker process, which initiates a new trial at this point.
 
@@
 
-The parallel asynchronous algorithm described above is based on the sequential information global search algorithm.
+The parallel asynchronous algorithm described above is based on the serial information global search algorithm.
 The theoretical substantiation of the algorithm convergence is given in \cite{Strongin2000}.
-Also, the synchronous parallelization schemes used earlier in solving a number of applied problems \cite{Kalyulin2017,Modorskii2016} are presented here.
-The novelty of the present work consists in a practical implementation and application of the asynchronous parallelization scheme featured by a higher efficiency in solving the problems with different computation costs of performing the trials at different points of the search domain.
-It was confirmed by the results of experiments described in the next section.
+The synchronous parallelization schemes used earlier in solving a number of applied problems \cite{Kalyulin2017,Modorskii2016} are also presented here.
+The novelty of the present work lies in the practical implementation and application of the asynchronous parallelization scheme featured by a higher efficiency in solving problems with different computational costs for performing trials at different points of the search domain.
+It was confirmed by the results of the experiments described in the next section.
 
@@
 
-According to the problem statements (\ref{eq:objective_func1}) and (\ref{eq:objective_func2}), the corresponding calculations were carried out in this work.
+According to problem statements (\ref{eq:objective_func1}) and (\ref{eq:objective_func2}), the corresponding calculations were carried out in this work.
 %ННГУ
-To carry out the numerical experiments, UNN supercomputer Lobachevsky was used (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node).
+The UNN supercomputer ``Lobachevsky'' (CentOS 7.2, SLURM, two CPUs Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node) was used for numerical experiments.
 The asynchronous global optimization algorithm was implemented using C++ (GCC 5.5.0 and Intel MPI were used); the objective function values were computed using Python 3.9.
@@
 First, the problem of searching for activation energies and pre-exponential multipliers of all reaction stages was solved.
-The number of optimized parameters is 30, which is two parameters for each of the fifteen reaction stages.
-For activation energies, the search range was set to $0 \leq E_i \leq 100$ kJ/mol, and $0 \leq E_i \leq 10^{12}$ for pre-exponential multipliers based on physicochemical considerations.
-However, based on the stiffness of the system of differential equations (\ref{eq:one}) and (\ref{eq:twelve}), no solution was found in such a wide range, this is due to the too high degree of pre-exponential multipliers included in the Arrhenius equation.
-Therefore, during the calculations, the value of the upper bound of the pre-exponential multipliers was reduced and the solution was obtained at the upper bounds of $10^5$.
+The number of optimized parameters is 30, i.e., two parameters for each of the fifteen reaction stages.
+For the activation energies, the search range was set to $0 \leq E_i \leq 100$ kJ/mol, and $0 \leq E_i \leq 10^{12}$ for the pre-exponential multipliers based on physicochemical considerations.
+However, based on the stiffness of the system of differential equations (\ref{eq:one}) and (\ref{eq:twelve}), no solution was found in such a wide range; this is due to the too high degree of pre-exponential multipliers included in the Arrhenius equation.
+Therefore, during the calculations, the value of the upper bound of the pre-exponential multipliers was reduced, and the solution was obtained at the upper bounds of $10^5$.
 The kinetic parameters found are presented in Table~\ref{table_res1}.
@@
 \begin{table}
-\caption{Calculated rate constants of reaction}
+\caption{Calculated rate constants of the reaction}
 \label{table_res1}
@@
 %ННГУ
-To evaluate the efficiency of the implemented optimization method, the results obtained using the sequential algorithm, the parallel synchronous, and the parallel asynchronous ones with the use of 8 nodes when solving the problem in the above statement were compared.
-The minimum values of the objective function found by respective methods, the times of solving the problem (in hours), and the values of speedup in time are presented in Table \ref{table_30D}. 
-During the experiments the parameter of the method $r=3.0$ from (\ref{R}) and the accuracy  $\epsilon = 10^{-3}\left\|b-a\right\|$ in the termination condition were used.
-Once the global search method achieves the preset accuracy, the solution was refined by Hook-Jeeves local method \cite{HookJeeves} with the accuracy $\epsilon = 10^{-5}\left\|b-a\right\|$.
+To evaluate the efficiency of the implemented optimization method, the results obtained  were compared applying the serial algorithm, the parallel synchronous and parallel asynchronous algorithms with the use of 8 nodes when solving the problem in the above statement.
+The minimum values of the objective function found by the respective methods, the time of solving the problem (in hours), and the values of speedup in time are presented in Table \ref{table_30D}. 
+During the experiments, the parameter of the method $r=3.0$ from (\ref{R}) and the accuracy  $\epsilon = 10^{-3}\left\|b-a\right\|$ in the termination condition were used.
+Once the global search method achieved the preset accuracy, the solution was refined by the Hook-Jeeves local method \cite{HookJeeves} with the accuracy $\epsilon = 10^{-5}\left\|b-a\right\|$.
 
@@
 
-The results have shown that both parallel algorithms had demonstrated a moderate speedup but had found better solutions than the sequential method.
-At that, the asynchronous algorithm had found a better solution than synchronous one.
-This explains also a smaller speedup of the asynchronous algorithm as compared with the synchronous one since the asynchronous method had executed more trials in the course of search for the optimal kinetic parameters of the problem.
+The results show that both parallel algorithms demonstrated a moderate speedup, but found better solutions than the serial method.
+At that, the asynchronous algorithm found a better solution than the synchronous one.
+This also explains a smaller speedup of the asynchronous algorithm as compared with the synchronous one, since the asynchronous method executed more trials in the course of search for the optimal kinetic parameters of the problem.
 
@@
   \includegraphics[width=0.7\linewidth]{res1.jpg}
-  \caption{Concentration profiles of key components when calculating the activation energies and pre-exponential reaction multipliers according to the problem statement~\ref{eq:objective_func1}.
+  \caption{Concentration profiles of key components when calculating the activation energies and pre-exponential reaction multipliers according to problem statement~\ref{eq:objective_func1}.
 Temperature: 276.2 K.
@@
 
-Next, the problem of finding constants $k_i$ of the reaction stages included in the system (\ref{eq:one})--(\ref{eq:twelve}) was solved.
+Next, the problem of finding the constants $k_i$ of the reaction stages included in the system (\ref{eq:one})--(\ref{eq:twelve}) was solved.
 Velocity constants were calculated for each of the temperatures (Table \ref{table_res2}).
@@
 %ННГУ
-When solving this problem, the operation performances of the sequential algorithm, of the parallel synchronous, and of the parallel asynchronous ones were also compared.  
-The minimum values of the objective function found by respective methods, the time of solving the problem (in hours), and the speedup in time with the use of 8 nodes are presented in Table \ref{table_15D}.
+When solving this problem, the operation performances of the serial algorithm, of the parallel synchronous and parallel asynchronous algorithms were also compared.  
+The minimum values of the objective function found by the respective methods, the time of solving the problem (in hours), and the speedup in time with the use of 8 nodes are presented in Table \ref{table_15D}.
 All the parameters of the method were the same as in the previous run.
@@
 
-The results have shown that all algorithms had found good solutions (in the values of the objective function).
-At that, the asynchronous algorithm had demonstrated two times higher speedup than the synchronous one.
-A good speedup of the asynchronous algorithm remains at a greater number of nodes as well, the results of its work are presented in Table \ref{tab_parall}.
+The results show that all algorithms found good solutions (in the values of the objective function).
+At that, the asynchronous algorithm demonstrated twice as much speedup than the synchronous one.
+The good speedup of the asynchronous algorithm remains with a larger number of nodes; the results of its work are presented in Table \ref{tab_parall}.
 
-Note that when solving the global optimization problem, the number of iterations of the parallel algorithm (and hence its speedup) depends significantly on the estimation of the Lipschitz constant of the objective function.
+Note that when solving the global optimization problem, the number of iterations of the parallel algorithm (and hence its speedup) significantly depends on the estimation of the Lipschitz constant of the objective function.
 The constant is adaptively estimated during the work of the algorithm and may vary depending on the accumulated search information.
-With a correct estimate of the Lipschitz constant the method can converge to a global optimum point faster than in the case of its incorrect estimate.
-This explains the effect of superlinear speedup observed in experiments.
+With a correct estimate of the Lipschitz constant, the method can converge to the global optimum point faster than in the case of its incorrect estimate.
+This explains the effect of superlinear speedup observed in the experiments.
 
@@
 %Часть УГНТУ
-The direct problem of chemical kinetics was solved with the found rate constants, the results of comparison with experimental data are shown in Fig.
+The direct problem of chemical kinetics was solved with the found rate constants, the results of comparison with the experimental data are shown in Fig.
 \ref{fig:res2} (an experiment at a temperature of 276.2 K is presented).
@@
   \includegraphics[width=0.7\linewidth]{res2.jpg}
-  \caption{Concentration profiles of key components when calculating the rate constants according to the problem statement~\ref{eq:objective_func2}.
+  \caption{Concentration profiles of key components when calculating the rate constants according to problem statement~\ref{eq:objective_func2}.
 Temperature: 276.2 K.
@@
 After calculating the rate constants for different temperature values, it is possible to calculate the parameters $E_i$ and $k_i^0$.
-So, for example, for the first stage of the reaction $E_1 = 24.65$ kJ/mol, $k_1^0 = 6.85 \cdot 10^4$ min$^{-1}$, for the third stage $E_3 = 75.57$ kJ/mol, $k_3^0 = 2.62 \cdot 10^{14}$ kg $\cdot$mol$^{-1}\cdot$min$^{-1}$.
-However, for some stages, the parameter values $k^0$ were fairly high, namely for the second stage $k_2^0 = 8.52 \cdot 10^{25}$ min$^{-1}$, for the seventh stage $k_7^0 = 2.82 \cdot 10^{26}$ kg mol$^{-2}$ min$^{-2}$.
+Thus, for example, for the first stage of the reaction $E_1 = 24.65$ kJ/mol, $k_1^0 = 6.85 \cdot 10^4$ min$^{-1}$, for the third stage $E_3 = 75.57$ kJ/mol, $k_3^0 = 2.62 \cdot 10^{14}$ kg $\cdot$mol$^{-1}\cdot$min$^{-1}$.
+However, for some stages, the parameter values $k^0$ were fairly high: for the second stage $k_2^0 = 8.52 \cdot 10^{25}$ min$^{-1}$, for the seventh stage $k_7^0 = 2.82 \cdot 10^{26}$ kg mol$^{-2}$ min$^{-2}$.
 Therefore, in future works, the constants corresponding to these kinetic parameters will be recalculated.
@@
 The solution may change if, for example, additional experimental data are found.
-However, the algorithm proposed in the paper provides a single solution within the search range, and high values of pre-exponential multipliers explain the presence of complex stages, which in turn consist of several stages that can be refined by chemists.
+However, the algorithm proposed in the paper provides a single solution within the search range, and the high values of the pre-exponential multipliers explain the presence of complex stages, which, in turn, consist of several stages that can be refined by chemists.
+The traditional scheme of assessing the quality of the problem solution as a deviation of the found solution from the exact one is not applicable here since the exact solution is not known.
 
 \section{Conclusions and Future Work}
-Thus, the article describes the search for kinetic parameters of the industrial chemical reaction of isobutane alkylation with olefins in the presence of sulfuric acid.
-A search was made for the rate constants of all reaction stages, as well as activation energies and pre-exponential multipliers have been calculated.
+The article describes the search for the kinetic parameters of the industrial chemical reaction of isobutane alkylation with olefins in the presence of sulfuric acid.
+The search was carried out for the rate constants of all reaction stages, and the activation energies and pre-exponential multipliers were calculated.
 The optimization method allowed us to find a fairly accurate description of the experimental data.
-In the future, it is planned to use this method and parallelization to eliminate high values of the pre-exponential reaction multipliers and to search for optimal conditions for the alkylation reaction using the developed kinetic model.
+In the future, it is planned to use this method and parallelization to eliminate the high values of the pre-exponential reaction multipliers and search for optimal conditions for the alkylation reaction using the developed kinetic model.
 
@@
 
-\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204, and by RFBR, project No.\,19-37-60014.
+\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204, and the Russian Foundation for Basic Research, project No.\,19-37-60014.
 

### 2022-05-05 2022/Sensors/kNN.tex
add one more paper
add response to comment 1 of reviewer 1

@@
 
-%%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
-%\usepackage{inputenc}
-%\usepackage[T2A,T1]{fontenc}
-%\usepackage[english,russian]{babel}
-%\usepackage{cmap}
-%%%%
+%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
+\usepackage{inputenc}
+\usepackage[T2A,T1]{fontenc}
+\usepackage[english,russian]{babel}
+\usepackage{cmap}
+%%%
 
@@
 To run Algorithms A and B, we need to compute the preimages $\{x_i^0, x_i^1, ..., x_i^L\}$ on one-dimensional scales for all objects $\omega_i$, and order them in ascending order.
-It is convenient to use RB-tree-based data structures to store ordered one-dimensional values.
-The number of instances of tree structures is equal to the number of scales.
+It is convenient to use RB-tree-based data structures to store ordered one-dimensional values, <Сысоев>алгоритмическая сложность операции поиска в котором имеет оценку O(logN).
+В качестве реализации RB-tree мы использовали контейнер set из стандартной библиотеки языка С++.
+<Сысоев> The number of instances of tree structures is equal to the number of scales.
 

### 2022-05-05 2022/Sensors/kNN.tex
first comment

--- 2022-04-15 2022/Sensors/kNN.tex
@@
 Such curves are used, for example, in global optimization \cite{Lera2015,Lera2018,Strongin2018,Gergel2021}, in numerical approximation of solutions to systems of nonlinear inequalities \cite{Lera2021}, in image processing \cite{Liang2008,Costa2012,Herrero2015}, etc.
+
+%Баркалов
+Отметим, что кроме space-filling curves cуществуют и другие схожие кривые, например, area-filling curves \cite{Nasso2022}.
+Однако широкое практическое применение получили именно Peano-Hilbert curves в силуv относительной простоты построения и наличия ряда полезных свойств, основные из которых будут отмечены в Section \ref{Peano} и Section \ref{MPeano}.
 

### 2022-05-06 2022/Sensors/kNN.tex
second comment

--- 2022-05-05 2022/Sensors/kNN.tex
@@
 
+%Баркалов
+Отметим, что кроме space-filling curves cуществуют и другие схожие кривые, например, area-filling curves \cite{Nasso2022}.
+Однако широкое практическое применение получили именно Peano-Hilbert curves в силу относительной простоты построения и наличия ряда полезных свойств, основные из которых будут отмечены в Section \ref{Peano} и Section \ref{MPeano}.
+
 An algorithm based on data curves can offer some advantages over tree-based algorithms by reducing the search time.
@@
 Note that the centre  $y^c(x)$ of the $M$-th partition subcube containing point $y(x)$ can be regarded as an approximation to $y(x)$ with the accuracy of $\varepsilon = 2^{-(M+1)}$, i.e.
-\begin{equation}
+\begin{equation}\label{prec}
 \max_{1 \leq j \leq N} \left| y_j^c(x)-y_j(x) \right| \leq \varepsilon = 2^{-(M+1)}.
@@
 A constructive way of establishing such a correspondence is described and theoretically justified in \cite{Strongin2000, Sergeyev2013}.
+%Баркалов
+Алгоритмическая сложность вычисления значения функции $y^c(x)$ зависит от размерности $N$, от уровня разбиения $M$ и составляется величину $O(M\cdot N)$.
 
@@
 Consider a family of Peano curves
-\begin{equation}
+\begin{equation}\label{EvSet}
 Y_L(x)=\{ y^0(x), y^1(x), ..., y^L(x) \}
@@
 \] 
+%Баркалов
+Алгоритмическая сложность вычисления всех прообразов $x^l \in [0, 1], \; 0 \leq l \leq L,$ зависит от размерности $N$, от уровня разбиения $M$, числа отображений $L$ и составляется величину $O(M\cdot N \cdot L)$.
 
@@
 \textbf{Theorem 1.} Let $y^*$ be an arbitrary point from the domain $D$ belonging to the interval with endpoints $y^\prime, y^{\prime\prime} \in D$, which differ in the values of the only coordinate, and let
-\[
+\begin{equation}\label{y_prime}
 |y_j^\prime - y^{\prime\prime}_j| \leq 2^{-p}, \; y^\prime_i = y^{\prime\prime}_i = y_i^*, \; 1 \leq i \leq N, i \ne j,
-\] 
+\end{equation} 
 where $p$ is an integer, $1 \leq p \leq L-1$, and $j$ is the number of coordinate whose values for points $y^*, y^\prime, y^{\prime\prime}$ are different.
@@
 \switchcolumn
+
+%Баркалов
+Как было сказано выше, вместо теоретической Hilbert curve $y(x)$ на практике используется evolvent $y^c(x)$, которая является приближением к $y(x)$ с точностью порядка $2^{-(M+1)}$, определяемой соотношением (\ref{prec}).
+При этом параметр $M$ может выбираться для каждой решаемой задачи индивидуально, исходя из точности представления исходных данных.
+Например, данные в  DS-Skin dataset (Section \ref{RealData}) представляют собой координаты пикселей изображения с разрешением 256 на 256.
+При отображении данных на стандартный гиперкуб $D$ from (\ref{D}) расстояние по координате между двумя точками не может быть меньше $1/256$.
+Использование параметра $M=10$ для построения evolvent будет достаточным в этом случае.
+В случае же присвоения численных значений небольшому числу categorical features (как это было сделано для CarEvaluation dataset, see Section \ref{RealData}) расстояние по координате между объектами будет большое, значение параметра $M$ можно снизить (в экспериментах использовалось значение $M=7$).
+
+Максимальное число одновременно используемых отображений в множестве $Y_L(x)$ from (\ref{EvSet}) также будет определяться точностью исходных данных, от которой зависит точность построения разверток $y^i(x)$.
+Будем счистать, что каждое из отображений $y^i(x)$ построено с использованием одного и того же значения $M$.
+Тогда значение $L$ должно удовлетворять неравенству $L \leq M$.
+В противном случае точки $y' = y^l(x'),\; y''=y^l(x'')$ из условия (\ref{y_prime}) будут соответствовать центру одного и того же подкуба при $p>M$, т.е.
+будет выполняться равенство $y' = y''$  что приведет к нарушению условий Теоремы 1.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Investigation of neighbour search times on real-world data}
+\subsection{Investigation of neighbour search times on real-world data}\label{RealData}
 

### 2022-05-06 2022/Sensors/kNN.tex
add response to comment 4 of reviewer 2

@@
 
+%Сысоев
+Следующий эксперимент проводился для определения зависимости времени поиска $K = 50$ соседей с ростом числа сэмплов для выбранной размерности $N = 4$.
+Число сэмплов менялось от 200 тыс.
+до 1 млн.
+Результаты, представленные на рис.XX, показывают что соотношение между процедурой А и процедурой B остается примерно тем же, а выигрыш процедуры А у kNN-ME method составляет порядка 10 раз.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-05-09 2022/Sensors/kNN.tex
Update kNN.tex
change comments

@@
 To run Algorithms A and B, we need to compute the preimages $\{x_i^0, x_i^1, ..., x_i^L\}$ on one-dimensional scales for all objects $\omega_i$, and order them in ascending order.
-It is convenient to use RB-tree-based data structures to store ordered one-dimensional values, <Сысоев>алгоритмическая сложность операции поиска в котором имеет оценку O(logN).
-В качестве реализации RB-tree мы использовали контейнер set из стандартной библиотеки языка С++.
-<Сысоев> The number of instances of tree structures is equal to the number of scales.
+It is convenient to use RB-tree-based data structures to store ordered one-dimensional values,
+%Сысоев
+алгоритмическая сложность операции поиска в котором имеет оценку $O(log(n))$. В качестве реализации RB-tree мы использовали контейнер set из стандартной библиотеки языка С++.
+The number of instances of tree structures is equal to the number of scales.
 
@@
 до 1 млн.
-Результаты, представленные на рис.XX, показывают что соотношение между процедурой А и процедурой B остается примерно тем же, а выигрыш процедуры А у kNN-ME method составляет порядка 10 раз.
+Результаты, представленные на Fig.~\ref{figN}, показывают что соотношение между процедурой А и процедурой B остается примерно тем же, а выигрыш процедуры А у kNN-ME method составляет порядка 10 раз.
+
+%Штанюк
+\begin{figure}
+\widefigure
+\includegraphics[width=0.5\linewidth]{figN.pdf}
+\caption{Running times of the algorithms with $N = 4$, $K = 50$\label{figN}}
+\end{figure}
 

### 2022-05-09 2022/Sensors/kNN.tex
Categorial features comments

@@
 
+%Штанюк
+Метод, основанный на пространственных кривых, может применятся при работе с категориальными переменными.
+Как известно, категориальные элементы образуют множество дискретных значений.
+Необходимо сопоставить каждому такому значению отображение на числовой отрезок, в нашем случае на [0,1].
+Расположение точек-отображений целесообразно на максимальном расстоянии друг от друга, в этом случае плотность заполняющих пространство кривых может быть уменьшена при сохранении точности обнаружения соседей.
+
+Для иллюстрации этого подхода был взят CarEvaluation dataset(Section \ref{RealData}), который включет в себя несколько категориальных атрибутов.
+Эти атрибуты носят названия buying, maint, doors, persons, lug_boot, safety, каждому из которых соответствуют 3-4 значения \cite{ds-2}.
+
+Отображение на числовой отрезок может быть сделано с помощью равномерной сетки.
+Отрезок делится на фиксированное количество равных частей, равное числу значений признаков (например 3 или 4, как в примере с CarEvaluation dataset).
+Далее, каждому значению переменной выбирается значение либо в середине соответствующей части отрезка, либо на одном из его концов.
+Если необходимо, то проводится округление значения до ближайшего разряда.
+
 %Баркалов

### 2022-05-09 2022/Sensors/kNN.tex
move comment

--- 2022-05-09 2022/Sensors/kNN.tex
@@
 %Баркалов
-Отметим, что кроме space-filling curves cуществуют и другие схожие кривые, например, area-filling curves \cite{Nasso2022}.
+Отметим тажке, что кроме space-filling curves cуществуют и другие схожие кривые, например, area-filling curves \cite{Nasso2022}.
 Однако широкое практическое применение получили именно Peano-Hilbert curves в силу относительной простоты построения и наличия ряда полезных свойств, основные из которых будут отмечены в Section \ref{Peano} и Section \ref{MPeano}.
@@
 For example, one possible implementation of searching for $K$ nearest neighbours could be as follows: in relation to $x^*$, we look for the nearest values to the left and right on the one-dimensional scale, choose the minimum one and continue searching sideways until exactly $K$ neighbours are found.
+
+%Штанюк
+Отметим, что метод, основанный на использовании space-filling curves, может применяться не только при работа с числовыми, но и при работе с категориальными features.
+Значения категориальных признаков являются дискретными, поэтому для применения данного подхода требуется сопоставить каждому дискретному значению численную величину, например, из интервала [0,1].
+Конкретный способ сопоставления численных значений категориальным переменным приведен в Section \ref{RealData} (в  CarEvaluation dataset).
+%Расположение точек на отрезке [0,1] целесообразно на максимальном расстоянии друг от друга, в этом случае плотность заполняющих пространство кривых может быть уменьшена при сохранении точности обнаружения соседей.
 
@@
 
+%Баркалов
+Также отметим, что данный dataset включет в себя несколько категориальных атрибутов, каждому из которых соответствуют 3-4 различных значения.
+В этом случае требуется выполнить сопоставление дискретным значениями численных характеристик.
+В качетсве численных значений при приведении эксперимента были выбраны внутренние узлы равномерной сетки, покрывающей отрезок [0,1]. 
+Число узлов сетки выбиралось равным числу значений признаков (3 или 4, соответствевнно). Далее, i-му значению категориального признака ставилось в соответствие значение в i-м узле сетки.
+
 \begin{specialtable}[H] 

### 2022-05-10 2022/Sensors/kNN.tex
red color for comments

--- 2022-05-09 2022/Sensors/kNN.tex
@@
 %Баркалов
-Отметим, что кроме space-filling curves cуществуют и другие схожие кривые, например, area-filling curves \cite{Nasso2022}.
-Однако широкое практическое применение получили именно Peano-Hilbert curves в силу относительной простоты построения и наличия ряда полезных свойств, основные из которых будут отмечены в Section \ref{Peano} и Section \ref{MPeano}.
+\textcolor[rgb]{1,0,0}{Отметим тажке, что кроме space-filling curves cуществуют и другие схожие кривые, например, area-filling curves \cite{Nasso2022}.
+Однако широкое практическое применение получили именно Peano-Hilbert curves в силу относительной простоты построения и наличия ряда полезных свойств, основные из которых будут отмечены в Section \ref{Peano} и Section \ref{MPeano}.}
 
@@
 %Баркалов
-Алгоритмическая сложность вычисления значения функции $y^c(x)$ зависит от размерности $N$, от уровня разбиения $M$ и составляется величину $O(M\cdot N)$.
+\textcolor[rgb]{1,0,0}{Алгоритмическая сложность вычисления значения функции $y^c(x)$ зависит от размерности $N$, от уровня разбиения $M$ и составляется величину $O(M\cdot N)$.}
 
@@
 For example, one possible implementation of searching for $K$ nearest neighbours could be as follows: in relation to $x^*$, we look for the nearest values to the left and right on the one-dimensional scale, choose the minimum one and continue searching sideways until exactly $K$ neighbours are found.
+
+%Штанюк
+\textcolor[rgb]{1,0,0}{Отметим, что метод, основанный на использовании space-filling curves, может применяться не только при работа с числовыми, но и при работе с категориальными features.
+Значения категориальных признаков являются дискретными, поэтому для применения данного подхода требуется сопоставить каждому дискретному значению численную величину, например, из интервала [0,1].
+Конкретный способ сопоставления численных значений категориальным переменным приведен в Section \ref{RealData} (в  CarEvaluation dataset).
+}%Расположение точек на отрезке [0,1] целесообразно на максимальном расстоянии друг от друга, в этом случае плотность заполняющих пространство кривых может быть уменьшена при сохранении точности обнаружения соседей.
 
@@
 %Баркалов
-Алгоритмическая сложность вычисления всех прообразов $x^l \in [0, 1], \; 0 \leq l \leq L,$ зависит от размерности $N$, от уровня разбиения $M$, числа отображений $L$ и составляется величину $O(M\cdot N \cdot L)$.
+\textcolor[rgb]{1,0,0}{Алгоритмическая сложность вычисления всех прообразов $x^l \in [0, 1], \; 0 \leq l \leq L,$ зависит от размерности $N$, от уровня разбиения $M$, числа отображений $L$ и составляется величину $O(M\cdot N \cdot L)$.}
 
@@
 %Баркалов
-Как было сказано выше, вместо теоретической Hilbert curve $y(x)$ на практике используется evolvent $y^c(x)$, которая является приближением к $y(x)$ с точностью порядка $2^{-(M+1)}$, определяемой соотношением (\ref{prec}).
+\textcolor[rgb]{1,0,0}{Как было сказано выше, вместо теоретической Hilbert curve $y(x)$ на практике используется evolvent $y^c(x)$, которая является приближением к $y(x)$ с точностью порядка $2^{-(M+1)}$, определяемой соотношением (\ref{prec}).
 При этом параметр $M$ может выбираться для каждой решаемой задачи индивидуально, исходя из точности представления исходных данных.
@@
 В случае же присвоения численных значений небольшому числу categorical features (как это было сделано для CarEvaluation dataset, see Section \ref{RealData}) расстояние по координате между объектами будет большое, значение параметра $M$ можно снизить (в экспериментах использовалось значение $M=7$).
-
 Максимальное число одновременно используемых отображений в множестве $Y_L(x)$ from (\ref{EvSet}) также будет определяться точностью исходных данных, от которой зависит точность построения разверток $y^i(x)$.
@@
 В противном случае точки $y' = y^l(x'),\; y''=y^l(x'')$ из условия (\ref{y_prime}) будут соответствовать центру одного и того же подкуба при $p>M$, т.е.
-будет выполняться равенство $y' = y''$  что приведет к нарушению условий Теоремы 1.
+будет выполняться равенство $y' = y''$  что приведет к нарушению условий Теоремы 1.}
 
@@
 %Сысоев
-алгоритмическая сложность операции поиска в котором имеет оценку $O(log(n))$. В качестве реализации RB-tree мы использовали контейнер set из стандартной библиотеки языка С++.
+\textcolor[rgb]{1,0,0}{алгоритмическая сложность операции поиска в котором имеет оценку $O(log(n))$. В качестве реализации RB-tree мы использовали контейнер set из стандартной библиотеки языка С++.}
 The number of instances of tree structures is equal to the number of scales.
@@
 %Сысоев
-Следующий эксперимент проводился для определения зависимости времени поиска $K = 50$ соседей с ростом числа сэмплов для выбранной размерности $N = 4$.
+\textcolor[rgb]{1,0,0}{Следующий эксперимент проводился для определения зависимости времени поиска $K = 50$ соседей с ростом числа сэмплов для выбранной размерности $N = 4$.
 Число сэмплов менялось от 200 тыс.
 до 1 млн.
-Результаты, представленные на Fig.~\ref{figN}, показывают что соотношение между процедурой А и процедурой B остается примерно тем же, а выигрыш процедуры А у kNN-ME method составляет порядка 10 раз.
+Результаты, представленные на Fig.~\ref{figN}, показывают что соотношение между процедурой А и процедурой B остается примерно тем же, а выигрыш процедуры А у kNN-KD method составляет порядка 10 раз.}
 
@@
 One particular feature of the dataset is the fact that the number of objects belonging to each of the 4 classes differs widely (Table \ref{tab:t2}).
+
+%Баркалов
+\textcolor[rgb]{1,0,0}{Также отметим, что данный dataset включет в себя несколько категориальных атрибутов, каждому из которых соответствуют 3-4 различных значения.
+В этом случае требуется выполнить сопоставление дискретным значениями численных характеристик.
+В качеcтве численных значений при приведении эксперимента были выбраны внутренние узлы равномерной сетки, покрывающей отрезок [0,1]. 
+Число узлов сетки выбиралось равным числу значений признаков (3 или 4, соответствевнно). Далее, i-му значению категориального признака ставилось в соответствие значение в i-м узле сетки. }
 

### 2022-05-14 2022/LION/main.tex
change the size of the figure
delete one word

@@
 One of such problem classes to which ML can be applied is the class of multicriterial optimization (MCO) models.
-In this case, ML is used, as a rule, in combination with metaheuristic algorithms \cite{Ma2014,Subraveti2019,Talbi2015} that in the case of multiextremal criteria lose in efficiency to deterministic methods \cite{Kvasov2018,Sergeyev2018}.
+In this case, ML is used, as a rule, in combination with metaheuristic algorithms \cite{Ma2014,Subraveti2019,Talbi2015} that in the case of multiextremal criteria lose in efficiency in comparison with deterministic methods \cite{Kvasov2018,Sergeyev2018}.
 
@@
 In practice, various scalarization techniques are often used to find effective solutions \cite{Ehrgott2005,GergelKozinov2020,Marler2004,Pardalos2017}.
-The present study uses the minimax convolution of partial criteria that possesses good theoretical properties
+The present study uses the minimax convolution of partial criteria that possesses good theoretical properties and allows one to find effective points by solving the problem
 \begin{equation}
-\min{f(y)} \to \min_{y\in D}{\max_{1 \leq i \leq s}{(f_i (y) \lambda_i )}} ,
+\max_{1 \leq i \leq s}{ \left\{ \lambda_i f_i (y) \right\} } \rightarrow \min, \; y \in D ,
 \label{eq:4}
 \end{equation}
-where $\lambda_i \geq 0$, $1 \leq i \leq s$, $\sum_{i=1}^s{\lambda_i}=1$, are numerical indicators of the significance of each criterion.
+where $\lambda_i \geq 0$, $1 \leq i \leq s$, $\sum_{i=1}^s{\lambda_i}=1$, are numerical indicators (weights) of the significance of each criterion.
 
@@
 In the framework of the proposed approach the information-statistical theory of global search is used for solving the scalar optimization problems (\ref{eq:4}).
-For problems with several variables ($N>1$) the dimensionality reduction on the base of Peano-type mappings \cite{Sergeyev2013,Strongin2000}.
+For problems with several variables ($N>1$) the dimensionality reduction on the base of Peano-type mappings \cite{Sergeyev2013,Strongin2000} is applied.
 While solving a problem (\ref{eq:4}) the algorithm generates a sequence of trials where the term ``trial'' means evaluation of criteria at a point of the feasible domain $D$ and computation of the convolution (\ref{eq:4}) at this point.
@@
   \item Select the subinterval with the highest characteristic.
-  \item Place the point of new trial within the subinterval with the highest characteristic. 
+  \item Place a point of new trial within the subinterval with the highest characteristic. 
   \item Map the new trial point to the point $y \in D$ using Peano evolvent and compute the value of the function (\ref{eq:4}) at this point. 
@@
 
-The second improvement is connected with a new method for calculating the characteristics of subintervals formed during univariate optimization on the internal $[0,1]$.
+The second improvement is connected with a new method for calculating the characteristics of subintervals formed during univariate optimization on the interval $[0,1]$.
 
@@
 
-In the second group of experiments 100 bi-criteria MCO problems have been solved. As the criteria the multiextremal functions
+In the second group of experiments 100 bi-criteria MCO problems have been solved. As the criteria, the multiextremal functions
 
@@
 Each multicriterial problem was reduced to 50 scalar problems (\ref{eq:4}) with different weight coefficients $\lambda$, distributed uniformly.
-Other algorithm's parameters were the same.
+Other parameters of the algorithm were the same.
 The averaged results can be found in Table~\ref{tab:2}.
-\textcolor[rgb]{1,0,0}{Fig.~\ref{fig:2} additionally shows the average value of HV index change depending of the number of trials.}
+Fig.~\ref{fig:2} additionally shows the average value of HV index depending on the number of trials.
 
@@
 \center
-\includegraphics[width=0.7\textwidth]{fig2.png}
+\includegraphics[width=0.8\textwidth]{fig2.png}
 \caption{Efficiency of the multicriterial algorithms compared} \label{fig:2}
@@
 
-\textcolor[rgb]{1,0,0}{The results show that embedding the machine learning into the global search algorithm leads to }a significant reduction of the number of trials (almost 3 times) with maintaining close values of indicators HV and DU.
+The results show that embedding the machine learning into the global search algorithm leads to a significant reduction of the number of trials (almost 3 times) with maintaining close values of indicators HV and DU.
 

### 2022-05-14 2022/PAVT/PAVT2022.tex

@@
 
-Now let us assume that $k\geq 0$ trials (in particular, $k$ can be equal to 0) are completed, and worker processes perform trials at the points $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.
-
+Now let us assume that $k\geq 0$ trials (in particular, $k$ can be equal to 0) are completed, and worker processes perform trials at the points
+$$y(x^{k+1}), \; y(x^{k+2}), ...,\; y(x^{k+p}).$$ 
 Each worker process, having completed its trial at some point (without any loss of generality, let us assume this point to be $y(x^{k+1})$ corresponding to Process No 1), sends the trial result to the master process.

### 2022-05-14 2022/Sensors/kNN.tex

@@
 %Баркалов
-\textcolor[rgb]{1,0,0}{Отметим тажке, что кроме space-filling curves cуществуют и другие схожие кривые, например, area-filling curves \cite{Nasso2022}.
-Однако широкое практическое применение получили именно Peano-Hilbert curves в силу относительной простоты построения и наличия ряда полезных свойств, основные из которых будут отмечены в Section \ref{Peano} и Section \ref{MPeano}.}
+\textcolor[rgb]{1,0,0}{We should also note that in addition to space-filling curves, there are other similar curves, such as area-filling curves \cite{Nasso2022}.
+However, it is the Peano-Hilbert curves that are commonly used due to the relative simplicity of their construction and a number of useful properties, of which the key ones will be discussed in Section \ref{Peano} and Section \ref{MPeano}.}
 
@@
 %Баркалов
-\textcolor[rgb]{1,0,0}{Алгоритмическая сложность вычисления значения функции $y^c(x)$ зависит от размерности $N$, от уровня разбиения $M$ и составляется величину $O(M\cdot N)$.}
+\textcolor[rgb]{1,0,0}{The algorithmic complexity of computing the value of the function $y^c(x)$ depends on the dimensionality $N$, on the partitioning level $M$ and is $O(M\cdot N)$.}
 
@@
 %Штанюк
-\textcolor[rgb]{1,0,0}{Отметим, что метод, основанный на использовании space-filling curves, может применяться не только при работа с числовыми, но и при работе с категориальными features.
-Значения категориальных признаков являются дискретными, поэтому для применения данного подхода требуется сопоставить каждому дискретному значению численную величину, например, из интервала [0,1].
-Конкретный способ сопоставления численных значений категориальным переменным приведен в Section \ref{RealData} (в  CarEvaluation dataset).
+\textcolor[rgb]{1,0,0}{Note that the method based on space-filling curves can be applied not only to numerical features but also to categorical features.
+The values of categorical features are discrete, so this approach requires a numerical value, e.g.
+from the interval [0,1], to be assigned to each discrete value.
+A specific method of assigning numerical values to categorical variables is described in Section \ref{RealData} for CarEvaluation dataset.
 }%Расположение точек на отрезке [0,1] целесообразно на максимальном расстоянии друг от друга, в этом случае плотность заполняющих пространство кривых может быть уменьшена при сохранении точности обнаружения соседей.
@@
 %Баркалов
-\textcolor[rgb]{1,0,0}{Алгоритмическая сложность вычисления всех прообразов $x^l \in [0, 1], \; 0 \leq l \leq L,$ зависит от размерности $N$, от уровня разбиения $M$, числа отображений $L$ и составляется величину $O(M\cdot N \cdot L)$.}
+\textcolor[rgb]{1,0,0}{The algorithmic complexity of computing all preimages $x^l \in [0, 1], \; 0 \leq l \leq L,$ depends on the dimensionality $N$, on the partitioning level $M$, on the number of mappings $L$ and is $O(M\cdot N \cdot L)$.}
 
@@
 %Баркалов
-\textcolor[rgb]{1,0,0}{Как было сказано выше, вместо теоретической Hilbert curve $y(x)$ на практике используется evolvent $y^c(x)$, которая является приближением к $y(x)$ с точностью порядка $2^{-(M+1)}$, определяемой соотношением (\ref{prec}).
-При этом параметр $M$ может выбираться для каждой решаемой задачи индивидуально, исходя из точности представления исходных данных.
-Например, данные в  DS-Skin dataset (Section \ref{RealData}) представляют собой координаты пикселей изображения с разрешением 256 на 256.
-При отображении данных на стандартный гиперкуб $D$ from (\ref{D}) расстояние по координате между двумя точками не может быть меньше $1/256$.
-Использование параметра $M=10$ для построения evolvent будет достаточным в этом случае.
-В случае же присвоения численных значений небольшому числу categorical features (как это было сделано для CarEvaluation dataset, see Section \ref{RealData}) расстояние по координате между объектами будет большое, значение параметра $M$ можно снизить (в экспериментах использовалось значение $M=7$).
-Максимальное число одновременно используемых отображений в множестве $Y_L(x)$ from (\ref{EvSet}) также будет определяться точностью исходных данных, от которой зависит точность построения разверток $y^i(x)$.
-Будем счистать, что каждое из отображений $y^i(x)$ построено с использованием одного и того же значения $M$.
-Тогда значение $L$ должно удовлетворять неравенству $L \leq M$.
-В противном случае точки $y' = y^l(x'),\; y''=y^l(x'')$ из условия (\ref{y_prime}) будут соответствовать центру одного и того же подкуба при $p>M$, т.е.
-будет выполняться равенство $y' = y''$  что приведет к нарушению условий Теоремы 1.}
+\textcolor[rgb]{1,0,0}{As it was mentioned above, instead of the theoretical Hilbert curve $y(x)$, the evolvent $y^c(x)$, which is an approximation to $y(x)$ with an accuracy of the order $2^{-(M+1)}$, determined by the relation  (\ref{prec}), is used in practice.
+In this case, the parameter $M$ can be chosen individually for each problem to be solved, based on the accuracy of the input data representation.}
+
+\textcolor[rgb]{1,0,0}{For example, the data in the DS-Skin dataset (Section \ref{RealData}) are pixel coordinates of an image with a resolution of 256 by 256.
+When mapping data to the standard hypercube $D$ from (\ref{D}), the coordinate distance between two points cannot be less than $1/256$.
+In this case, using the parameter $M=10$ to construct the evolvent will be sufficient.
+If we assign numerical values to a small number of categorical features (as was done for the CarEvaluation dataset, see Section \ref{RealData}), the  coordinate distance between the features will be large, so the value of $M$ can be reduced (in our experiments, we used $M=7$).}
+
+\textcolor[rgb]{1,0,0}{The maximum number of simultaneous mappings in the set $Y_L(x)$ from (\ref{EvSet}) will also be determined by the accuracy of the input data, which affects the accuracy of constructing the evolvents $y^i(x)$.
+Let us assume that each of the mappings  $y^i(x)$ is constructed using the same value of $M$.
+Then the value of $L$ must satisfy the inequality $L \leq M$.
+Otherwise the points $y' = y^l(x'),\; y''=y^l(x'')$ from condition (\ref{y_prime}) will correspond to the centre of the same subcube with $p>M$, i.e.
+the equality  $y' = y''$ will be fulfilled, leading to violation of the conditions of Theorem 1.}
 
@@
 %Сысоев
-\textcolor[rgb]{1,0,0}{алгоритмическая сложность операции поиска в котором имеет оценку $O(log(n))$. В качестве реализации RB-tree мы использовали контейнер set из стандартной библиотеки языка С++.}
+\textcolor[rgb]{1,0,0}{where the algorithmic complexity of the search operation has the estimate $O(log(n))$.
+We used the container set from the standard C++ library as an implementation of the RB-tree.}
 The number of instances of tree structures is equal to the number of scales.
@@
 %Сысоев
-\textcolor[rgb]{1,0,0}{Следующий эксперимент проводился для определения зависимости времени поиска $K = 50$ соседей с ростом числа сэмплов для выбранной размерности $N = 4$.
-Число сэмплов менялось от 200 тыс.
-до 1 млн.
-Результаты, представленные на Fig.~\ref{figN}, показывают что соотношение между процедурой А и процедурой B остается примерно тем же, а выигрыш процедуры А у kNN-KD method составляет порядка 10 раз.}
+\textcolor[rgb]{1,0,0}{The following experiment was conducted to determine the dependence of the search time for $K = 50$ neighbours with increasing number of samples for the chosen dimensionality $N = 4$.
+The number of samples was varied from 200 thousand to 1 million.
+The results presented in Figure~\ref{figN} show that the difference between procedure A and procedure B remains approximately the same, and the gain of procedure A over the kNN-KD method is about 10 times.}
 
@@
 %Баркалов
-\textcolor[rgb]{1,0,0}{Также отметим, что данный dataset включет в себя несколько категориальных атрибутов, каждому из которых соответствуют 3-4 различных значения.
-В этом случае требуется выполнить сопоставление дискретным значениями численных характеристик.
-В качеcтве численных значений при приведении эксперимента были выбраны внутренние узлы равномерной сетки, покрывающей отрезок [0,1]. 
-Число узлов сетки выбиралось равным числу значений признаков (3 или 4, соответствевнно). Далее, i-му значению категориального признака ставилось в соответствие значение в i-м узле сетки. }
+\textcolor[rgb]{1,0,0}{Also note that this dataset will include several categorical attributes, each corresponding to 3-4 different values.
+In this case, it is necessary to assign numerical characteristics to discrete values.
+The numerical values used in the experiment were the internal nodes of a uniform grid covering the interval $[0,1]$. 
+The number of grid nodes was chosen to be equal to the number of feature values (3 or 4, respectively). Then, the $i$-th value of the categorical feature was assigned to the $i$-th node of the grid.}
 

### 2022-05-19 2022/Sensors/kNN.tex
add response to comment 2 of reviewer 1

@@
 
+%Сысоев
+\textcolor[rgb]{1,0,0}{Представленные выше результаты экспериментов показывают, что даже при числе сэмплов в несколько миллионов времена работы рассмотриваемых в статье алгоритмов отстаются небольшими (в пределах сотен микросекунд).
+Таким образом, вопросы распараллеливания и реализации многопоточных версий разработанных алгоритмов в статье не рассматриваются.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-05-19 2022/Sensors/kNN.tex
add comments and figure for WISDM dataset

@@
 \textbf{DS-CarEvaluation} -- a dataset of dimension 6, representing data from a car valuation model based on the values of 6 features \cite{ds-2}.
+
+%Штанюк
+\item
+\textbf{DS-WISDM} -- a dataset of dimension 3, representing data from Smartphone and Smartwatch Activity and Biometrics Dataset Data Set \cite{ds-3}.
+
 \end{enumerate}
@@
 We employed in our experiments the kNN-KD method from the scikit-learn library; this method was used with default settings.
-For our kNN-ME method, we used the evolvent density $m=12$, the number of evolvents simultaneously used was $L=10$ for datasets 1 -- 3 and $L=7$ for the last dataset.
+For our kNN-ME method, we used the evolvent density $m=12$, the number of evolvents simultaneously used was $L=10$ for datasets 1 -- 3 and $L=7$ for the dataset 4.
 
@@
 The results of the comparison on this dataset also demonstrate that the kNN-ME method is faster compared to the kNN-KD method.
+
+%Штанюк
+\textcolor[rgb]{1,0,0}{Представляет интерес изучение зависимости времени поиска от числа соседей при большом количестве объектов.
+Для такого исследования был выбран датасет DS-WISDM \cite{ds-3}.
+The raw accelerometer and gyroscope sensor data is collected from the smartphone and smartwatch at a rate of 20Hz.
+It is collected from 51 test subjects as they perform 18 activities for 3 minutes apiece.
+Традиционное использование этого датасета предполагает обработку raw данных, которая включает разбиение на блоки фиксированного размера и вычисление статистических характеристик для каждого блока.
+В дальнейшем, каждый блок рассматривается в качестве самостоятельного объекта с набором присущих ему характеристик, определяющих активность субъекта за рассматриваемый временной промежуток.
+В рамках нашего исследования мы не стали использовать этот традиционный подход и использовали только три пространственные координаты без привязки к времени.
+Перед проведением вычислительного эксперимента из набора данных были удалены дубликаты объектов.}
+
+%Штанюк
+\begin{figure}
+\widefigure
+\includegraphics[width=0.5\linewidth]{figW.pdf}
+\caption{Running times of the algorithms with $N = 3$, $7 million objects$\label{figW}}
+\end{figure}
+
+%Штанюк
+\textcolor[rgb]{1,0,0}{
+На рисунке \label{figW} приведены результаты моделирования для почти 7 миллионов объектов, относящихся к 18 классам.
+Из двух процедур нахождения соседей была оставлена kNN-ME(A) и замеры времени производились для этой процедуры с количеством разверток L = 8,9,10.
+При числе соседей K<100 процедура kNN-ME(A) работает быстрее, чем kNN-KD.
+}
 

### 2022-05-19 2022/Sensors/kNN.tex
last commemt in Russian

--- 2022-05-14 2022/Sensors/kNN.tex
@@
 \textbf{DS-CarEvaluation} -- a dataset of dimension 6, representing data from a car valuation model based on the values of 6 features \cite{ds-2}.
+
+%Штанюк
+\item
+\textcolor[rgb]{1,0,0}{\textbf{DS-WISDM} -- a dataset of dimension 3, representing data from Smartphone and Smartwatch Activity and Biometrics Dataset Data Set \cite{ds-3}.}
+
 \end{enumerate}
@@
 We employed in our experiments the kNN-KD method from the scikit-learn library; this method was used with default settings.
-For our kNN-ME method, we used the evolvent density $m=12$, the number of evolvents simultaneously used was $L=10$ for datasets 1 -- 3 and $L=7$ for the last dataset.
+For our kNN-ME method, we used the evolvent density $m=12$, the number of evolvents simultaneously used was $L=10$ for datasets 1 -- 3, $L=7$ for the dataset 4 and $L=8$ for the dataset 5.
 
@@
 \includegraphics[width=0.5\linewidth]{fig6a.pdf}
-\caption{Comparison of neighbour search times for the CarEvoluation dataset\label{fig6a}}
+\caption{Comparison of neighbour search times for the CarEvaluation dataset\label{fig6a}}
 \end{figure}
 The results of the comparison on this dataset also demonstrate that the kNN-ME method is faster compared to the kNN-KD method.
+
+%Штанюк
+\textcolor[rgb]{1,0,0}{Представляет интерес изучение зависимости времени поиска ближайших соседей от их числа при большом количестве объектов.
+Для такого исследования был выбран датасет DS-WISDM \cite{ds-3}.
+The raw accelerometer and gyroscope sensor data is collected from the smartphone and smartwatch at a rate of 20Hz.
+It is collected from 51 test subjects as they perform 18 activities for 3 minutes apiece.
+Традиционное использование этого датасета предполагает обработку raw data, которая включает разбиение данных на блоки фиксированного размера и вычисление статистических характеристик для каждого блока.
+В дальнейшем, каждый блок рассматривается в качестве самостоятельного объекта с набором присущих ему характеристик, определяющих активность субъекта за рассматриваемый временной промежуток.
+При этом число блоков будет существенно меньше исходного числа объектов.
+В рамках нашего исследования для сохранения большого числа объектов в датасете мы не стали использовать этот традиционный подход и работали с необработанными данными.
+Для определения ближайших соседей использовались только пространственные координаты, без привязки ко времени.
+Перед проведением вычислительного эксперимента из исходного датасета были удалены встречающиеся дубликаты.
+В получившемся наборе осталось порядка 7 млн.
+объектов, относящихся к 18 классам.
+На рисунке \label{figW} приведены результаты  поиска ближайших соседей в этом дататесе.
+Из двух процедур нахождения соседей была оставлена kNN-ME(A) как более эффективная по сравнению с kNN-ME(B); использовалось число evolvents L = 8.
+Как и в предыдущих экспериментах, результаты показывают, что при числе соседей K<100 процедура kNN-ME(A) также работает быстрее, чем kNN-KD.}
+
+%Штанюк
+\begin{figure}
+\widefigure
+\includegraphics[width=0.5\linewidth]{figW.pdf}
+\caption{Running times of the algorithms with $N = 3$, 7 million objects\label{figW}}
+\end{figure}
+
+%Сысоев
+\textcolor[rgb]{1,0,0}{Представленные выше результаты экспериментов показывают, что даже при числе сэмплов в несколько миллионов времена работы рассмотриваемых в статье алгоритмов отстаются небольшими (в пределах миллисекунды).
+Таким образом, вопросы распараллеливания и реализации многопоточных версий разработанных алгоритмов в статье не рассматриваются.}
 

### 2022-05-20 2022/Sensors/kNN.tex
before translation of the last comment

@@
 \item
-\textcolor[rgb]{1,0,0}{\textbf{DS-WISDM} -- a dataset of dimension 3, representing data from Smartphone and Smartwatch Activity and Biometrics Dataset Data Set \cite{ds-3}.}
+\textcolor[rgb]{0,1,0}{\textbf{DS-WISDM} -- a dataset of dimension 3, representing accelerometer and gyroscope time-series sensor data collected from a smartphone and smartwatch \cite{ds-3}.}
 
@@
 %Штанюк
-\textcolor[rgb]{1,0,0}{Представляет интерес изучение зависимости времени поиска ближайших соседей от их числа при большом количестве объектов.
+\textcolor[rgb]{0,1,0}{Представляет интерес изучение зависимости времени поиска ближайших соседей от их числа при большом количестве объектов.
 Для такого исследования был выбран датасет DS-WISDM \cite{ds-3}.
@@
 В дальнейшем, каждый блок рассматривается в качестве самостоятельного объекта с набором присущих ему характеристик, определяющих активность субъекта за рассматриваемый временной промежуток.
-При этом число блоков будет существенно меньше исходного числа объектов.
+После такой обработки число блоков будет существенно меньше исходного числа объектов.
 В рамках нашего исследования для сохранения большого числа объектов в датасете мы не стали использовать этот традиционный подход и работали с необработанными данными.
@@
 объектов, относящихся к 18 классам.
-На рисунке \label{figW} приведены результаты  поиска ближайших соседей в этом дататесе.
-Из двух процедур нахождения соседей была оставлена kNN-ME(A) как более эффективная по сравнению с kNN-ME(B); использовалось число evolvents L = 8.
-Как и в предыдущих экспериментах, результаты показывают, что при числе соседей K<100 процедура kNN-ME(A) также работает быстрее, чем kNN-KD.}
+На Figure \ref{figW} приведены результаты  поиска ближайших соседей в этом датасете.
+Из двух наших процедур нахождения соседей была оставлена kNN-ME(A) как более эффективная по сравнению с kNN-ME(B); при этом использовалось число evolvents L = 8.
+Как и в предыдущих экспериментах, результаты показывают, что при числе соседей K<100 процедура kNN-ME(A) также работает быстрее, чем the kNN-KD method.}
 
@@
 \includegraphics[width=0.5\linewidth]{figW.pdf}
-\caption{Running times of the algorithms with $N = 3$, 7 million objects\label{figW}}
+\caption{\textcolor[rgb]{0,1,0}{Running times of the algorithms with $N = 3$, 7 million objects}\label{figW}}
 \end{figure}
@@
 %Сысоев
-\textcolor[rgb]{1,0,0}{Представленные выше результаты экспериментов показывают, что даже при числе сэмплов в несколько миллионов времена работы рассмотриваемых в статье алгоритмов отстаются небольшими (в пределах миллисекунды).
-Таким образом, вопросы распараллеливания и реализации многопоточных версий разработанных алгоритмов в статье не рассматриваются.}
+\textcolor[rgb]{0,1,0}{Представленные выше результаты экспериментов показывают, что даже при числе объектов в несколько миллионов времена работы рассматриваемых в статье алгоритмов остаются небольшими (в пределах миллисекунды).
+Таким образом, вопросы распараллеливания и реализации многопоточных версий разработанных алгоритмов в статье не рассматривались.}
 

### 2022-05-21 2022/Sensors/kNN.tex
translate last comment

--- 2022-05-19 2022/Sensors/kNN.tex
@@
 %%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
-\usepackage{inputenc}
-\usepackage[T2A,T1]{fontenc}
-\usepackage[english,russian]{babel}
-\usepackage{cmap}
+%\usepackage{inputenc}
+%\usepackage[T2A,T1]{fontenc}
+%\usepackage[english,russian]{babel}
+%\usepackage{cmap}
 %%%
@@
 \item
-\textcolor[rgb]{1,0,0}{\textbf{DS-WISDM} -- a dataset of dimension 3, representing data from Smartphone and Smartwatch Activity and Biometrics Dataset Data Set \cite{ds-3}.}
+\textcolor[rgb]{1,0,0}{\textbf{DS-WISDM} -- a dataset of dimension 3, representing accelerometer and gyroscope time-series sensor data collected from a smartphone and a smartwatch \cite{ds-3}.}
 
@@
 %Штанюк
-\textcolor[rgb]{1,0,0}{Представляет интерес изучение зависимости времени поиска ближайших соседей от их числа при большом количестве объектов.
-Для такого исследования был выбран датасет DS-WISDM \cite{ds-3}.
+\textcolor[rgb]{1,0,0}{It is of interest to study the dependence of the nearest-neighbour search time on the number of such neighbours in the case of a large number of objects.
+For this study, we chose the DS-WISDM dataset \cite{ds-3}.
 The raw accelerometer and gyroscope sensor data is collected from the smartphone and smartwatch at a rate of 20Hz.
 It is collected from 51 test subjects as they perform 18 activities for 3 minutes apiece.
-Традиционное использование этого датасета предполагает обработку raw data, которая включает разбиение данных на блоки фиксированного размера и вычисление статистических характеристик для каждого блока.
-В дальнейшем, каждый блок рассматривается в качестве самостоятельного объекта с набором присущих ему характеристик, определяющих активность субъекта за рассматриваемый временной промежуток.
-При этом число блоков будет существенно меньше исходного числа объектов.
-В рамках нашего исследования для сохранения большого числа объектов в датасете мы не стали использовать этот традиционный подход и работали с необработанными данными.
-Для определения ближайших соседей использовались только пространственные координаты, без привязки ко времени.
-Перед проведением вычислительного эксперимента из исходного датасета были удалены встречающиеся дубликаты.
-В получившемся наборе осталось порядка 7 млн.
-объектов, относящихся к 18 классам.
-На рисунке \label{figW} приведены результаты  поиска ближайших соседей в этом дататесе.
-Из двух процедур нахождения соседей была оставлена kNN-ME(A) как более эффективная по сравнению с kNN-ME(B); использовалось число evolvents L = 8.
-Как и в предыдущих экспериментах, результаты показывают, что при числе соседей K<100 процедура kNN-ME(A) также работает быстрее, чем kNN-KD.}
+The raw accelerometer and gyroscope sensor data is collected from a smartphone and a smartwatch at a rate of 20Hz.
+It is collected from 51 test subjects as they perform 18 activities for 3 minutes apiece.
+Traditional use of this dataset involves raw data processing, which includes splitting the data into blocks of fixed size and calculating statistical characteristics for each block.
+Subsequently, each block is treated as an independent object with a set of inherent characteristics determining the subject's activity during a given time interval.
+After such processing, the number of blocks will be substantially smaller than the original number of objects.}
+
+\textcolor[rgb]{1,0,0}{In our study, in order to keep a large number of objects in the dataset, we did not use this traditional approach and worked with raw data.
+Only spatial coordinates, with no time reference, were used to determine the nearest neighbours.
+Before conducting the computational experiment, we removed the duplicates found in the original dataset.
+After that, about 7 million objects belonging to 18 classes remained in the resulting set.
+Figure \ref{figW} shows the results of the nearest neighbour search in this dataset.
+Of our two neighbour finding procedures, kNN-ME(A) was retained as a more efficient one compared to kNN-ME(B); the number of evolvents $L = 8$ was used in this case.
+Similar to previous experiments, the results show that when the number of neighbours $K$ is less than $100$, the kNN-ME(A) procedure is also faster compared to the kNN-KD method.}
 
@@
 \includegraphics[width=0.5\linewidth]{figW.pdf}
-\caption{Running times of the algorithms with $N = 3$, 7 million objects\label{figW}}
+\caption{\textcolor[rgb]{1,0,0}{Running times of the algorithms with $N = 3$, 7 million objects}\label{figW}}
 \end{figure}
@@
 %Сысоев
-\textcolor[rgb]{1,0,0}{Представленные выше результаты экспериментов показывают, что даже при числе сэмплов в несколько миллионов времена работы рассмотриваемых в статье алгоритмов отстаются небольшими (в пределах миллисекунды).
-Таким образом, вопросы распараллеливания и реализации многопоточных версий разработанных алгоритмов в статье не рассматриваются.}
+\textcolor[rgb]{1,0,0}{The experimental results presented above show that even with the number of objects in millions the running times of the algorithms considered in the paper remain small (within milliseconds).
+Therefore, the issues of parallelizing and implementing multi-threaded versions of the developed algorithms have not been considered in this paper.}
 

### 2022-05-23 2022/Sensors/kNN.tex
revised paper submitted

@@
 Let us assume that each of the mappings  $y^i(x)$ is constructed using the same value of $M$.
-Then the value of $L$ must satisfy the inequality $L \leq M$.
-Otherwise the points $y' = y^l(x'),\; y''=y^l(x'')$ from condition (\ref{y_prime}) will correspond to the centre of the same subcube with $p>M$, i.e.
+Then the value of $L$ must satisfy the inequality $L < M$.
+Otherwise the points $y' = y^l(x'),\; y''=y^l(x'')$ from condition (\ref{y_prime}) will correspond to the centre of the same subcube with $p \geq M$, i.e.
 the equality  $y' = y''$ will be fulfilled, leading to violation of the conditions of Theorem 1.}
@@
 \includegraphics[width=0.5\linewidth]{figW.pdf}
-\caption{\textcolor[rgb]{1,0,0}{Running times of the algorithms with $N = 3$, 7 million objects}\label{figW}}
+\caption{Running times of the algorithms with $N = 3$, 7 million objects\label{figW}}
 \end{figure}

### 2022-06-02 2022/ICNAAM/ICNAAM.tex
template for ICNAAM paper

--- 2021-07-07 2021/ICNAAM/2021_ICNAAM.tex
@@
 \usepackage{graphicx}
-\usepackage{tikz}
-\usetikzlibrary{shapes.geometric}
-\usetikzlibrary{positioning}
-\usetikzlibrary{patterns}
-\usetikzlibrary{arrows,decorations.markings}
-\usetikzlibrary{spy}
 
@@
 % Title portion
-\title{Simulation and Optimization of Injector Parameters with OpenFOAM and Globalizer Tools}
+\title{Title}
 
@@
 \author[aff1]{Ilya Lebedev}
-\author[aff2]{Daniil Ryazanov}
-\author[aff2]{Sergey Strijak}
 
 \affil[aff1]{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia}
-\affil[aff2]{Ivannikov Institute for System Programming of the RAS, Moscow, Russia}
 \corresp[cor1]{Corresponding author: konstantin.barkalov@itmm.unn.ru}
@@
 \begin{abstract}
-The paper presents the results of solving the problem of simulation and research of processes taking place in a fuel system injector.
-Simulation has been carried using the tools of the OpenFOAM system.
-The problem of differential pressure minimization at the injector inlet and outlet was set and solved.
-This problem was a Lipschitz optimization problem with a black-box type objective function.
-To solve it, we used an efficient global search algorithm implemented in the Globalizer system.
+The paper presents results ...
 \end{abstract}
@@
 
-As engineering objects and conditions of their functioning become more complicated, it results in a significant increase in the complexity of corresponding mathematical models and numerical methods for their analysis.
-Nowadays, the main (and often the only possible) tool for such analysis is supercomputer simulation of the object's behavior.
-For this purpose, CAD/CAM/CAE-systems are widely used.
-A well-known example of open source software of this kind is OpenFOAM, an open platform for numerical simulation of continuum mechanics problems \cite{OpenFOAM}).
-
-In this study, we used the OpenFOAM system to simulate a fuel injector.
-In the simulation process, the pressure drop at the inlet and outlet of the injector was calculated.
-The problem of minimizing the pressure drop in order to save fuel by reducing the concentration of combustible substances in the mixture after leaving the ejector was set and solved.
-The parameters used in the problem included the width of the gap between the injector inlet and the injector outlet.
-
 \section{GLOBAL SEARCH ALGORITHM}
 
-The global optimization problem
-\begin{eqnarray}\label{problem}
-&\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
-&D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\},\label{D}
-\end{eqnarray}
-can serve as a mathematical model for the process of choosing optimum values of parameters $y=(y_1,y_2,...,y_N)$ of the engineering object under consideration.
-The choice of the value  $y^*$ that determines the best option  is reduced to an approximate (with some required accuracy) solution of the above optimization problem based on the evaluation of object characteristics for the set of options $y^i\in D, \; 1 \leq i \leq k$, determined by available resources.
-
-The optimal parameter selection problems under consideration are characterized by the fact that the objective function $\varphi(y)$ is not defined analytically; there is only some algorithm to calculate its values at the points of the domain $D$.
-In this case, one search trial (calculating the function value in a particular point of the search domain) is a computationally intensive operation \cite{Kalyulin2017}.
-
-Multi-extremal optimization problems are essentially more time-consuming in comparison with other types of optimization problems because global optimum is an integral characteristic of the problem to be solved and demands investigation of the whole search domain.
-As a result, the search for the global optimum is reduced to generating  some covering (grid) in the parameter domain and choosing the best function value on this grid.
-The amount of calculations can be reduced by creating a non-uniform coverage of the search domain: the grid has to be dense enough in the vicinity of the global optimum and more sparse farther away from the solution being sought.
-
-A typical assumption used in many global optimization methods \cite{Evtushenko2013,Jones2009,Zilinskas2010,Pinter1996} is the assumption that the objective function $\varphi(y)$ satisfies the Lipschitz condition
-\[
-\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D, 0<L<\infty.
-\]
-
-At the same time, various dimensionality reduction schemes are widely used to reduce the complexity of designing global optimization algorithms.
-By means of  such schemes, it is possible to reduce solution of multidimensional optimization problems to the solution of one or more nested one-dimensional optimization problems.
-
-One approach to dimensionality reduction is to use Peano-type space filling curves (\textit{evolvents}).
-Using evolvents $y(x)$, which continuously and uniquely map a unit interval onto a hypercube, reduces the problem of finding the minimum of the function $\varphi(y)$ in the domain $D$ to finding the minimum of a one-dimensional function  on the interval $[0,1]$
-\[
-\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}}.
-\]
-
-The above assumptions and approaches have led to the development of a number of global search algorithms based on the information-statistical approach \cite{Strongin2000}.
-These algorithms are implemented in the Globalizer software system for solving global optimization problems \cite{globalizerSystem}.
-Here, we present a scheme of the basic global search algorithm (GSA).
-This is a Divide-the-Best algorithm, i.e.
-it is aimed to conduct a subsequent search trial in the most promising (from the point of view of global minimum detection) search subdomain.
-
-The first two trials are performed at the boundary points of the search domain $x^0 = 0$ and $x^1 = 1$.
-The selection of the point  $x^{k+1}, \; k \geq 1,$  of any next trial is carried out by the following steps.
-
-Step 1. Renumber the points  $x^1,...,x^k$ of the preceding trials by the lower indices in increasing order of the coordinate values, i.e.
-\begin{equation}\label{x_i}
-0=x_0<x_1<\dots <x_k<x_{k+1}=1,
-\end{equation}
-and juxtapose to them the values $z_i=\varphi(y(x_i)), \; 0 \leq i\leq k+1,$  computed at these points.
-
-Step 2. For each interval $(x_{i-1},x_i), \; 1\leq i \leq k,$ compute the characteristics $R(i)$ using some formulae.
-
-Step 3. Find the interval $(x_{t-1},x_t),$ with the maximum characteristic 
-\[
-R(t)=\max\{R(i):1 \leq i \leq k+1\}.
-\]
-
-Step 4. Execute the next trial in the inner point of the interval  $(x_{t-1},x_t)$, i.e. $x^{k+1} \in (x_{t-1},x_t)$.
-
-Step 5. Check termination condition  $\left(x_t-x_{t-1}\right)^{1/N}<\epsilon$, where $t$   is the number of the interval with the maximum characteristic and $\epsilon>0$ is the predefined accuracy.
-
-A detailed description of the algorithm, its various modifications (including parallel ones), and the corresponding convergence  theory are presented in \cite{Strongin2000,Barkalov2010,Strongin2020}.
+The global optimization problem \cite{Strongin2000}
 
@@
 
-To carry out numerical experiments, we used the OpenFOAM software package \cite{OpenFOAM}.
-It is a set of tools for representing, discretizing and manipulating hydrodynamic fields.
-OpenFOAM also provides methods for solving systems of algebraic equations.
-The OpenFOAM serves to perform a complete cycle of numerical solution of continuum mechanics problems using the finite volume method: from the mathematical formulation by setting initial conditions, boundary conditions, computational geometry, and computational grid to a representation of the results in the form of hydrodynamic fields.
+%\begin{figure}%[ht]
+%\includegraphics[width=1.0\linewidth]{fig1.png}
+%\caption{Schematic representation of the computational domain, with the geometry parameterization points marked in red and the black points independent of the optimization iterations.
+The grey lines show the possible deformation of the geometry.}
+%\label{fig}
+%\end{figure}
 
-In our particular case, each iteration of the optimal solution search process was a separate numerical calculation of the hydrodynamic fields.
-The injector simulation was divided into several key steps:
-\begin{enumerate}
-\item Mathematical formulation.
-This includes a procedure for selecting equations describing the dynamics of a fluid or gas, setting boundary conditions and initial conditions.
-It does not change from iteration to iteration.
-\item Parameterization. Selection of varying parameters that will vary from iteration to iteration.
-\item Generating the geometry and the computational grid. In line with Parameterization, a procedure for renewing the geometry and computational grid is performed.
-\item Calculation of hydrodynamic fields. At each iteration, the system of equations describing fluid dynamics is solved numerically.
-\item Interpretation. Physical interpretation of the results after the optimization procedure.
-\end{enumerate}
-
-The fluid dynamics for the problem described in this paper is defined by the Navier-Stokes equations for an incompressible fluid:
-\[
-\nabla \cdot \vec{U} = 0,
-\]
-\[
-\frac{\partial \vec{U}}{\partial t} + \nabla \cdot \left( \vec{U} \otimes \vec{U}\right) - \nabla \cdot \nu \left( \nabla \vec{U} + (\nabla \vec{U})^T\right) = - \frac{1}{\rho} \nabla p,
-\]
-where $\vec{U}$ is the velocity vector, $\nu$  is the liquid viscosity (assumed constant), $p$  is the overpressure, $\rho$ is the density of the liquid.
-In this system of equations, there are two unknown variables: $\vec{U}$  and $p$.
-For each of these variables, initial conditions are set:
-\[
-\vec{U}(x,y) = 0,
-\]
-\[
-p(x,y) = 0.
-\]
-There are three types of boundaries in the computational domain; special boundary conditions are set on each of them. 
-On a rigid wall:
-\[
-\nabla p \cdot \vec{n} = 0,
-\]
-\[
-\vec{U} = 0.
-\]
-At the entrance to the computational domain:
-\[
-\nabla p \cdot \vec{n} = 0,
-\]
-\[
-\vec{U} = \vec{U_i}.
-\]
-At the exit of the computational domain:
-\[
- p = 0,
-\]
-\[
-\vec{U} = \vec{U_o}.
-\]
-
-The geometric representation of the computational domain is shown in Figure 1, with the parameterization points highlighted in red.
-The geometry is parametrized by increasing the distance between the red points symmetrically with respect to the channel centreline.
-
-\begin{figure}%[ht]
-%\includegraphics[width=1.0\linewidth]{fig1.png}
-
-    \centering
-    \begin{tikzpicture}[scale=0.5,>=stealth]
-         \draw (-5,-2.5,0)--(-15,-2.5,0)--(-15,2.5,0)--(-5,2.5,0.0);%--(0,0.5,0.0)--(5,2.5,0)--(15,2.5,0)--(15,-2.5,0);
-         \draw (-5,-2.5) .. controls (0,0) .. (5,-2.5);
-         \draw (5,2.5,0)--(15,2.5,0)--(15,-2.5,0)--(5,-2.5,0.0);
-         \draw (-5,2.5) .. controls (0,0) .. (5,2.5);
-
-         \filldraw [black] (-5,-2.5,0) circle (4pt);
-         \filldraw [black] (-15,-2.5,0) circle (4pt);
-         \filldraw [black] (-15,2.5,0) circle (4pt);
-         \filldraw [black] (-5,2.5,0.0) circle (4pt);
-
-         \filldraw [black] (5,2.5,0) circle (4pt);
-         \filldraw [black] (15,2.5,0) circle (4pt);
-         \filldraw [black] (15,-2.5,0) circle (4pt);
-         \filldraw [black] (5,-2.5,0.0) circle (4pt);
-
-         \filldraw [red] (0,0.6,0.0) circle (4pt);
-         \filldraw [red] (0,-0.6,0.0) circle (4pt);
-
-         \draw [->,ultra thick] (-16,0,0) --node[above right]{$x$} (-13,0,0);
-         \draw [->,ultra thick] (-16,0,0) --node[above left]{$y$} (-16,3,0);
-
-         \draw [<->, thick] (-15,-4,0)--node[below]{$6\;\;cm$} (15,-4,0);
-
-         \draw [dashed] (-15,-2.5) -- (-15,-5);
-         \draw [dashed] (15,-2.5) -- (15,-5);
-
-         \draw [dashed] (15,-2.5) -- (16.5,-2.5);
-         \draw [dashed] (15,2.5) -- (16.5,2.5);
-
-         \draw [<->, thick] (16,-2.5,0)--node[above, rotate=-90]{$0.5\;\;cm$} (16,2.5,0);
-
-         \draw[dashed] (-17,0,0)--(17,0,0);
-
-         \draw[opacity=0.2] (-5,2.5) .. controls (0,1) .. (5,2.5);
-         \draw[opacity=0.2] (-5,-2.5) .. controls (0,-1) .. (5,-2.5);
-
-         \filldraw [red,opacity=0.2] (0,1.4,0.0) circle (4pt);
-         \filldraw [red,opacity=0.2] (0,-1.4,0.0) circle (4pt);
-
-         \draw[opacity=0.2] (-5,2.5) .. controls (0,2) .. (5,2.5);
-         \draw[opacity=0.2] (-5,-2.5) .. controls (0,-2) .. (5,-2.5);
-
-         \filldraw [red,opacity=0.2] (0,2.15,0.0) circle (4pt);
-         \filldraw [red,opacity=0.2] (0,-2.15,0.0) circle (4pt);
-
-         \draw[opacity=0.2] (-5,2.5) .. controls (0,3) .. (5,2.5);
-         \draw[opacity=0.2] (-5,-2.5) .. controls (0,-3) .. (5,-2.5);
-
-         \filldraw [red,opacity=0.2] (0,2.9,0.0) circle (4pt);
-         \filldraw [red,opacity=0.2] (0,-2.9,0.0) circle (4pt);
-
-        \draw[opacity=0.2] (-5,2.5) .. controls (0,4) .. (5,2.5);
-        \draw[opacity=0.2] (-5,-2.5) .. controls (0,-4) .. (5,-2.5);
-
-        \filldraw [red,opacity=0.2] (0,3.65,0.0) circle (4pt);
-        \filldraw [red,opacity=0.2] (0,-3.65,0.0) circle (4pt);
-
-    \end{tikzpicture}
-\caption{Schematic representation of the computational domain, with the geometry parameterization points marked in red and the black points independent of the optimization iterations.
-The grey lines show the possible deformation of the geometry.}
-\label{fig}
-\end{figure}
-
-The calculation was carried out with the standard pisoFoam solver and the PISO algorithm \cite{Issa1986_2,Issa1986_1}.
-A total of 300 finite elements were involved in the calculation.
-The process of finding the effective channel thickness took 3395 seconds on an Intel(R) Xeon(R) CPU X5670 2.93 GHz, during which time 69 iterations of the global search algorithm were performed.
-The horizontal component of the velocity at the entrance to the computational domain $\vec{U_i}$  was 10 m/s.
-The pressure at the exit of the computational domain $p_o$ was equal to zero, just like all components of the velocity vector at the exit $\vec{U_o}$ .
-Linear dimensions of the computational domain are shown in the schematic of Figure 1.
-As a result of solving the problem, the optimum channel thickness (0.59 cm) was found, giving the minimum fuel consumption.
+The calculation was carried out ...
 

### 2022-06-10 2022/RSD/paper.tex
after review

@@
 The trials are performed at the boundary points $x^0 = 0$ and $x^1 = 1$ as well as at $(p-2)$ arbitrary different internal points $x^2, x^3, ..., x^{p-1}$ in the interval $[0,1]$.
+After executing the first iteration of the method, i.e. at $n=1$, the number of trials will be $k=p$.
+%After the first iteration, i.e. at $n=1$, the number of trials will be $k=p$.
 
@@
 
-The theoretical basis of the presented algorithm (denoted hereafter as \textit{global search algorithm} or GSA) see in more details in \cite{Strongin2000}.
+The theoretical basis of the presented algorithm (denoted hereafter as \textit{global search algorithm} or GSA) see in more details in \cite{Strongin2000,Strongin2018}.
 Note briefly that the characteristics of the intervals (\ref{R}) used in the algorithm can be considered as a quantitative measure how particular interval is promising from the viewpoint of finding the global minimum inside it.

### 2022-06-13 2022/ICNAAM/ICNAAM.tex
first version of introduction

@@
 \usepackage{graphicx}
+
+%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
+\usepackage{inputenc}
+\usepackage[T2A,T1]{fontenc}
+\usepackage[english,russian]{babel}
+\usepackage{cmap}
+%%
 
@@
 \author[aff1]{Konstantin Barkalov\corref{cor1}}
-\author[aff1]{Ilya Lebedev}
+\author[aff1]{Marina Usova}
 
@@
 
-The problem of designing new engineering objects is inseparably connected with the problem of selecting the best values of item parameters, because the choice of specific values of parameters (with a fixed overall structure) can significantly affect its characteristics.
+В настоящее время методы глобальной оптимизации используются для решения широкого круга задач, возникающих в различных областях науки и техники.
+
+Например, традиционной сферой применения методов глобальной оптимизации является идентификация значений параметров математических моделей по данным экспериментов.
+К их числу относятся, например, inverse problems of chemical kinetics.
+В задачах такого вида требуется провести поиск значений неизвестных параметров модели, при которых результаты моделирования наиболее близки к результатам, полученным экспериментально.
+
+Число параметров, которые требуется идентифицировать подобным образом, для задач химической кинетики может составлять десятки и сотни.
+Использование детерминированных методов глобальной оптимизации для решения задачах такой размерности крайне ограничено из-за чрезвычайно больших вычислительных затрат на покрытие search domain точками испытаний.
+Это остается справедливым даже в случае использования эффективных алгоритмов (например, \cite{Paulavicius2011,Evtushenko2009,Jones2009}), строящих существенно неравномерные покрытия.
+
+Многие обратные задачи характеризуются тем, что зависимость от разных групп параметров носит разный характер.
+Например, от одной группы параметров зависимость может быть близка к линейной, тогда как по второй группе зависимость может носить сложный многоэкстремальный характер.
+
+В этом случае решение задачи можно организовать по схеме recursive optimization.
+Решение  многоэкстремальной подзадачи, для которой требуется использовать сложные алгоритмы глобальной оптимизации, будет проводиться на верхнем уровне рекурсии.
+Унимодальные подзадачи (каждая из которых соответствует фиксированному набору значений первой части параметров) будут решаться на нижнем уровне.
+Для их решения можно применять обладающие быстрой сходимостью методы локальной оптимизации.
+
+Однако заранее указать разделение на группы параметров с разным характером поведения целевой функции, как правило, не представляется возможным, т.к.
+целевая функция в обратных задачах задается как черный ящик.
+Таким образом, актуальной является разработка алгоритмов минимизации существенно многомерных функций, в которых учитывается разный характер зависимости целевой функции от разных групп параметров. 
+В данной работе предложен конкретный механизм такого разделения, основанный на <указать>.
+В качестве примеров, подтверждающих работоспособность предложенной схемы, решены как тестовые, так и прикладные задачи.
 

### 2022-06-16 2022/ICNAAM/ICNAAM.tex
draft of the paper in Russian

@@
 % Title portion
-\title{Title}
+\title{Использование методов машинного обучения для разделения параметров в задачах глобальной оптимизации большой размерности}
 
@@
 
-\section{GLOBAL SEARCH ALGORITHM}
+\section{ОПТИМИЗАЦИОННЫЕ АЛГОРИТМЫ И РЕДУКИЦЯ РАЗМЕРНОСТИ}
 
-The global optimization problem \cite{Strongin2000}
+Задачи оптимизации: глобальные и локальные
+
+Многошаговая схема
+
+Используемые методы глобальной и локальной оптимизации
+
+\section{РАЗДЕЛЕНИЕ ПАРАМЕТРОВ ЗАДАЧИ}
+
+Описать алгоритм разделения параметров и формирования подзадач для многошаговой схемы
 
@@
 
-The calculations were performed on the computer cluster of the Institute for Systems Programming and the UniHub web-lab \cite{UniHub}.
-The web-lab architecture is based on the cloud computing model where resources (servers, networks, storage systems, applications, etc.) are provided remotely as a set of different-level services with on-demand servicing.
+Описание тестовых задач
+
+Описание результатов экспериментов
 

### 2022-06-17 2022/ICNAAM/ICNAAM.tex
Section 2 in Russian

@@
 В настоящее время методы глобальной оптимизации используются для решения широкого круга задач, возникающих в различных областях науки и техники. 
-
 Например, традиционной сферой применения методов глобальной оптимизации является идентификация значений параметров математических моделей по данным экспериментов.
@@
 
-Задачи оптимизации: глобальные и локальные
+Рассмотрим задачу оптимизации вида
+\begin{eqnarray}\label{main_problem}
+& \varphi(y^\ast)=\min{\left\{\varphi(y): y\in D\right\}}, \nonumber \\
+& D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
+\end{eqnarray}
+Будем предполагать, что функция $\varphi(y)$ является многоэкстремальной и удовлетворяет Lipschitz condition
+\[
+\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in D,\; 0<L<\infty,
+\]
+with the constant $L$ unknown a priori.
+Одновременно с этим будет предполагать, что зависимость целевой функции от $N_1$ параметров является многоэкстремальной, а от оставшихся $N_2 = N - N_1$ параметров зависимость являеся близкой к линейной.
+При этом разделение  параметров является неизвестным.
+Такой характер зависимостей является типичным, например, для обратных задач химической кинетики [ссылка].
+В данной работе предлагается подход к разделению параметров задачи на указанные две группы, конкретный механизм описан в следуюем разделе.
 
-Многошаговая схема
+Учет подобной особенности задач может существенно снизить вычислительную сложность поиска оптимума.
+В самом деле, в соответствии с известной схемой рекурсивной оптимизации \cite{Carr} решение исходной задачи можно проводить следующим образом:
+\begin{equation}\label{global_problem}
+\varphi(y^*) = \min_{y\in D} \varphi (y) = \min_{y^1\in D_1} f(y^1), \; D_1=\left\{y\in R^{N_1}: a_i\leq y_i \leq b_i, \; 1\leq i \leq N_1\right\},
+\end{equation}
+где 
+\begin{equation}\label{local_problem}
+f(y^1) = \min_{ y^2 \in D_2} \varphi(y^1,y^2), \; D_2=\left\{y\in R^{N_2}: a_i\leq y_i \leq b_i, \; N_1+1\leq i \leq N\right\}.
+\end{equation}
+В соответствии с (\ref{global_problem}) вычисление одного значения функции $\varphi^1 (y^1)$ (данный процесс будем называть search trial) подразумевает решение унимодальной задачи (\ref{local_problem}), которое может быть выполнено одним из методов локальной оптимизации.
+Эффективность и проработанность методов локальной оптимизации позволяет решать подзадачи (\ref{local_problem}) с точностью, значительно превышающей точность методов глобальной оптимизации.
+Соответственно, можно рассматривать задачу (\ref{global_problem}) как задачу глобальной оптимизации, в которой значение функции может быть вычислено с высокой точностью, но данная операция является трудоемкой.
 
-Используемые методы глобальной и локальной оптимизации
+В проведенном исследовании для решения задач глобальной оптимизации (\ref{global_problem}) мы использовали global search algorithm в сочетании со схемой редукции размерности на основе кривых Пеано [ссылки].
+Данный алгорим зарекомендовал себя как мощный инструмент для решения сложных многоэкстремальных задач [ссылки].
+Для решения локальной задачи (\ref{local_problem}) мы использовали Hooke-Jeevse pattern search method (see, e.g., \cite{Kelley}).
+В отличие от методов типа BFGS (see, e.g., \cite{Nocedal}), данный алгоритм не требует вычисления градиента и хорошо работает в условиях наличия вычислительных ошибок в значениях функций, что является типичным при решении прикладных задач.
 

### 2022-06-22 2022/ICNAAM/ICNAAM.tex

@@
 \begin{abstract}
-The paper presents results ...
+В работе представлены результаты исследования подхода к решению задач оптимизации, в которых зависимость от разных групп параметров носит разный характер.
+Предложена схема выделения параметров задачи, которые оказывают локальное влияние на целевую функцию, что позволяет решать существенно многомерные задачи с использованием nested optimization scheme.
+При этом на разных уровнях рекурсии используются разные оптимизационные алгоритмы.
+Предложенный подход показал свою работоспособность при решении тестовых и прикладных задач.
+
 \end{abstract}
@@
 Например, традиционной сферой применения методов глобальной оптимизации является идентификация значений параметров математических моделей по данным экспериментов.
-К их числу относятся, например, inverse problems of chemical kinetics.
+К их числу относятся, например, inverse problems of chemical kinetics \cite{Gubaydullin2021}.
 В задачах такого вида требуется провести поиск значений неизвестных параметров модели, при которых результаты моделирования наиболее близки к результатам, полученным экспериментально.
@@
 
-В проведенном исследовании для решения задач глобальной оптимизации (\ref{global_problem}) мы использовали global search algorithm в сочетании со схемой редукции размерности на основе кривых Пеано [ссылки].
-Данный алгорим зарекомендовал себя как мощный инструмент для решения сложных многоэкстремальных задач [ссылки].
+В проведенном исследовании для решения задач глобальной оптимизации (\ref{global_problem}) мы использовали global search algorithm в сочетании со схемой редукции размерности на основе кривых Пеано \cite{Sergeyev2013,Lera2021}.
+Данный алгорим зарекомендовал себя как мощный инструмент для решения сложных многоэкстремальных задач \cite{Kvasov2013,Kalyulin2017,Cavoretto2021}.
 Для решения локальной задачи (\ref{local_problem}) мы использовали Hooke-Jeevse pattern search method (see, e.g., \cite{Kelley}).

### 2022-06-22 2022/ICNAAM_NetOpt/My theses.tex
Первая версия тезисов по "Использование аппроксимации нейронными сетями при решении задач глобальной оптимизации"

% !!!!!!!!!!!!!!!!!!   ВНИМАНИЕ !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
% Заголовки разделов формируются при помощи команд \section{}, \subsection{}, \subsubsection{}
% Не используйте уровень вложенности заголовков больше трех!
% -----------------------------
% Для оформления теорем, лемм, следствий используйте окружения 
% Def     - Определение
% Teor    - Теорема
% Lem     - Лемма
% Predl   - Предложение
% Ass     - Утверждение
% Cor     - Следствие
% Example - Пример
% -----------------------------
% Доказательство теоремы начинается командой \proof и завершается командой \endproof
% -----------------------------
% Литература помещается в окружение biblio.

\documentclass[11pt, oneside, a4paper]{article}
%\usepackage[cp1251]{inputenc} % кодировка
\usepackage[utf8]{inputenc} % кодировка
\usepackage[english, russian]{babel} % Русские и английские переносы
\usepackage{graphicx}          % для включения графических изображений
\usepackage[unicode, pdftex]{hyperref}
\usepackage{cite}              % для корректного оформления литературы
\usepackage{enumitem}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{mathtext}

%стилевой пакет
\usepackage{schoolseminar2022}

%\renewcommand{\thefootnote}{*}\footnote{Работа выполнена при частичной финансовой поддержке .....}

\begin{document}
% \udk     - универсальный десятичный классификатор
% \msc     - Индекс предметной классификации (Mathematics Subject Classification)
% \title   - название статьи
% \authors - список авторов
\setcounter{page}{1}

\udk{КодУДК}

\title{Использование аппроксимации нейронными сетями при решении задач глобальной оптимизации \footnote{Работа выполнена при частичной финансовой поддержке .....}}

\authors{Баркалов К.А., Лебедев И.Г., Надумин Д.В.}
\organizations{ННГУ им. Н.И. Лобачевского\\
Институт информационных технологий математики и механики}

% \section{название} - заголовок раздела первого уровня
% \subsection{название} - заголовок раздела второго уровня
% \subsubsection{название} - заголовок раздела третьего уровня

%\bigskip

В работе рассматривается задача поиска глобального минимума двумерной многоэкстремальной функции, с использованием аппроксимации целевой функции нейронной сетью. \cite{fio_bib1}\cite{fio_bib2}\cite{fio_bib3}\cite{fio_bib4}

В работах \cite{fio_bib5},\cite{fio_bib6} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п. В данной работе проводились эксперименты с серией двухмерных функциий из генератора GKLS.

Задача экспериментов состояла в том, чтобы проверить возможность уменьшения количества вычислений функции для её оптимизации. На приведённых ниже рисунках изображены исследуемые функции, слева исходная функция, а справа аппроксимация этой функции полученная с помощью нейронной сети.

\begin{figure}[h]
	\begin{center}
			\includegraphics[scale=0.25]{figure/approximate}
			\includegraphics[scale=0.25]{figure/true}			
			\caption{Пример аппроксимированной (слева) и истинной функции из GKLS} %% подпись к рисунку
            \label{fio_ris1}
	\end{center}
\end{figure}

Для аппроксимации была выбрана нейронная сеть прямого распространения сигнала с 5-ю слоями по 32 нейрона в каждом. В качестве активационной функции в каждом нейроне была выбрана логистическая функция.

Для обучения нейронной сети использовался заранее подготовленный набор из 160 точек испытания распределенных в области поиска.
Значения функции переводились в нормальную шкалу перед обучением. Аппроксимация также в нормальной шкале от 0 до 1.

После аппроксимации модель(нейронная сеть) передавалась в метод глобальной оптимизации.
В качестве такого метода был выбран метод имитации отжига\cite{fio_bib7}.

Программная реализация аппроксимации функции с помощью нейронной сети проводится в среде Jupyter Notebook – сервис для интерактивных вычислений. В качестве языка программирования выбран Python, который удобен для математических вычислений. Используются следующие библиотеки языка Python: Numpy, Matplotlib, SciPy, Keras.

Алгоритм решения задачи оптимизации заключается в последовательном вызове алгоритма оптимизации на аппроксимированной функции и аппроксимации функции нейронной сетью \cite{fio_bib8}. В качестве критерия остановки использовалась близость найденной точки к априори известному минимуму.

В результате проведенных исследований удалось сократить число вычислений целевой функции у 80\% задач.

\begin{biblio}

\bibitem{fio_bib1} Cybenko, G. V. Approximation by Superpositions of a Sigmoidal function. Mathematics of Control Signals and Systems. — 1989. — Т. 2, № 4. — С. 303—314.

\bibitem{fio_bib2} Warren S. McCulloch, Walter Pitts // California State University. – Режим доступа:https://home.csulb.edu/~cwallis/382/readings/482/mccolloch.logical.calculus.ideas.1943.pdf

\bibitem{fio_bib3} Paul Werbos // National Science Foundation. – Режим доступа: \url{https://www.researchgate.net/publication/35657389_Beyond_regression_new_tools_for_prediction_and_analysis_in_the_behavioral_sciences}

\bibitem{fio_bib4} Y.LeCun, B.Boser // ATI\&T Bell Laboratories. – Режим доступа: \url{http://yann.lecun.com/exdb/publis/pdf/lecun-89e.pdf}

\bibitem{fio_bib5} Sergeyev Y.D., Kvasov D.E. Diagonal'nyye metody global'noy optimizatsii. Fizmatlit, 2008. P. 352.

\bibitem{fio_bib6} Gaviano M., Lera D., Kvasov D. E., Sergeyev Y. D. Software for generation of classes of test functions with known local and global minima for global optimization // ACM Transactions on Mathematical Software. 2003. Vol. 29. P. 469-480.

\bibitem{fio_bib7} Kirkpatrick, S.; Gelatt Jr,C.D.;Vecchi, M. P. (1983). "Optimization by Simulated Annealing". Science. 220 (4598): 671–680.

\bibitem{fio_bib8} Мордухович Б. Ш. Методы аппроксимаций в задачах оптимизации и управления / Мордухович Б. Ш. . – М.: «Наука», 1988. – 360 c.

\end{biblio}

\msc{Код MSC2010}

\end{document}

### 2022-06-22 2022/ICNAAM_NetOpt/My theses.tex
Небольшое изменение
some corrections

@@
 
-\authors{Баркалов К.А., Лебедев И.Г., Надумин Д.В.}
+\authors{Карпенко С.Н., Лебедев И.Г., Надумин Д.В.}
 \organizations{ННГУ им. Н.И. Лобачевского\\
@@
 
-В работе рассматривается задача поиска глобального минимума двумерной многоэкстремальной функции, с использованием аппроксимации целевой функции нейронной сетью.
-\cite{fio_bib1}\cite{fio_bib2}\cite{fio_bib3}\cite{fio_bib4}
+В работе рассматривается задача поиска глобального минимума многоэкстремальной функции.
+Для ускорения процесса поиска минимума используется аппроксимации целевой функции нейросетью \cite{fio_bib1,fio_bib2,fio_bib3,fio_bib4}.
 
-В работах \cite{fio_bib5},\cite{fio_bib6} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.
-В данной работе проводились эксперименты с серией двухмерных функциий из генератора GKLS.
+В работах \cite{fio_bib5,fio_bib6} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.
+В данной работе проводились эксперименты с серией двумерных функций из генератора GKLS.
 
-Задача экспериментов состояла в том, чтобы проверить возможность уменьшения количества вычислений функции для её оптимизации.
-На приведённых ниже рисунках изображены исследуемые функции, слева исходная функция, а справа аппроксимация этой функции полученная с помощью нейронной сети.
+Цель эксперимента состояла в том, чтобы проверить возможность уменьшения количества вычислений целевой функции при решении задачи оптимизации.
+На рис.~\ref{fio_ris1} изображены исследуемые функции, слева исходная функция, а справа аппроксимация этой функции полученная с помощью нейронной сети.
 
@@
 			\includegraphics[scale=0.25]{figure/true}			
-			\caption{Пример аппроксимированной (слева) и истинной функции из GKLS} %% подпись к рисунку
+			\caption{Пример аппроксимированной (слева) и истинной (справа) функции из класса GKLS} %% подпись к рисунку
             \label{fio_ris1}
@@
 
-После аппроксимации модель(нейронная сеть) передавалась в метод глобальной оптимизации.
-В качестве такого метода был выбран метод имитации отжига\cite{fio_bib7}.
+После аппроксимации модель (нейронная сеть) передавалась в метод глобальной оптимизации.
+В качестве такого метода использовался метод имитации отжига\cite{fio_bib7}.
 
-Программная реализация аппроксимации функции с помощью нейронной сети проводится в среде Jupyter Notebook – сервис для интерактивных вычислений.
-В качестве языка программирования выбран Python, который удобен для математических вычислений.
-Используются следующие библиотеки языка Python: Numpy, Matplotlib, SciPy, Keras.
+Программная реализация аппроксимации функции с помощью нейронной сети проведена в среде Jupyter Notebook (сервис для интерактивных вычислений).
+В качестве языка программирования использован Python, который удобен для математических вычислений.
+Были задействованы следующие библиотеки языка Python: Numpy, Matplotlib, SciPy, Keras.
 
-Алгоритм решения задачи оптимизации заключается в последовательном вызове алгоритма оптимизации на аппроксимированной функции и аппроксимации функции нейронной сетью \cite{fio_bib8}.
+Алгоритм решения задачи оптимизации заключался в последовательном вызове алгоритма оптимизации на аппроксимированной функции и аппроксимации функции нейронной сетью \cite{fio_bib8}.
 В качестве критерия остановки использовалась близость найденной точки к априори известному минимуму.
 
-В результате проведенных исследований удалось сократить число вычислений целевой функции у 80\% задач.
+Как показали проведенные эксперименты, использование нейросетевой аппроксимации позволяет сократить число вычислений целевой функции в 80\% задач.
 
@@
 
-\bibitem{fio_bib1} Cybenko, G. V. Approximation by Superpositions of a Sigmoidal function. Mathematics of Control Signals and Systems. — 1989. — Т. 2, № 4. — С. 303—314.
+\bibitem{fio_bib1} Cybenko, G. V. Approximation by Superpositions of a Sigmoidal function. Mathematics of Control Signals and Systems. — 1989. — Т/. 2, № 4. — С. 303—314.
 

### 2022-06-22 2022/ICNAAM_NetOpt/My theses.tex
one more correction

@@
 
-\title{Использование аппроксимации нейронными сетями при решении задач глобальной оптимизации \footnote{Работа выполнена при частичной финансовой поддержке .....}}
+\title{Использование аппроксимации нейронными сетями при решении задач глобальной оптимизации\footnote{Работа выполнена при поддержке Министерства науки и высшего образования РФ (проект \textnumero~0729-2020-0055) и научно-образовательного математического центра <<Математика технологий будущего>> (проект \textnumero~075-02-2021-1394).}}
 
@@
 Цель эксперимента состояла в том, чтобы проверить возможность уменьшения количества вычислений целевой функции при решении задачи оптимизации.
-На рис.~\ref{fio_ris1} изображены исследуемые функции, слева исходная функция, а справа аппроксимация этой функции полученная с помощью нейронной сети.
+На рис.~\ref{fio_ris1} изображены исследуемые функции: слева исходная функция, а справа аппроксимация этой функции полученная с помощью нейронной сети.
 
@@
 Для обучения нейронной сети использовался заранее подготовленный набор из 160 точек испытания распределенных в области поиска.
-Значения функции переводились в нормальную шкалу перед обучением. Аппроксимация также в нормальной шкале от 0 до 1.
+Значения функции приводились к диапазону [0,1] перед обучением, аппроксимация также строилась в диапазоне [0,1].
 

### 2022-06-23 2022/ICNAAM/ICNAAM.tex
добавлены описание метода и результаты экспериментов

@@
 
-Описать алгоритм разделения параметров и формирования подзадач для многошаговой схемы
+Для разделения параметров задачи на локальные и глобальные предлагается следующий механизм действий:
+
+Шаг 1.
+Зафиксировать некоторое сечение $\widetilde{y} = (\widetilde{y}_1, \widetilde{y}_2, … ,\widetilde{y}_N)$ целевой функции $\varphi(y)$.
+Выбор сечения осуществляется одним из предложенных способов:
+\begin{enumerate}
+\item сечение фиксируется из физического смысла прикладной задачи;
+\item в качестве фиксируемого сечения принимается решение, найденное локальным методом за $K$ итераций, используемый локальный метод – Hooke-Jeevse pattern search method [используется по умолчанию].
+\end{enumerate}
+
+Шаг 2. Для очередной переменной $y_i$, $1\leq i \leq N$, провести исследование на локальность:
+\begin{enumerate}
+\item вычислить $P$ значений целевой функции $\varphi(y)$ в зафиксированном сечении $\widetilde{y}$ при равномерно изменяемой исследуемой переменной $y_i$ в своей области значений $a_i\leq y_i \leq b_i$;
+\item сформировать множество $Q$, представляющее собой набор из $P$ пар вида «значение исследуемой переменной – значение целевой функции в сечении $\widetilde{y}$ при указанном значении исследуемой переменной»
+\[
+Q = \left\{ (\widetilde{y}^k_i,  \varphi({\widetilde{y}^k})): \widetilde{y}^k = (\widetilde{y}_1, \widetilde{y}_2, … , \widetilde{y}^k_i, … ,\widetilde{y}_N), \widetilde{y}^k_i \sim U[a_i, b_i], 1\leq i \leq N, 1\leq k \leq P \right\};
+\]
+\item построить регрессионную модель заданной степени $deg$ на основе данных $Q$;
+\item вычислить оценку $R^2$ и провести классификацию переменной:
+\begin{itemize}
+\item если $R^2$ меньше порога вхождения в локальные функции $T$, то добавить переменную к множеству глобальных переменных,
+\item иначе – добавить переменную к множеству локальных переменных.
+\end{itemize}
+\end{enumerate}
+
+Шаг 3.
+Если локальные переменные были обнаружены – запустить последовательную блочную многошаговую схему, на верхнем уровне которой глобальные переменные, а на нижнем - локальные.
+В противном случае запустить обычный АГП, считая все переменные глобальными.
+
+Комментарий 1. В основе проводимой регрессии лежит метод наименьших квадратов (МНК).
+%
+%Комментарий 1. В основе проводимой регрессии лежит метод наименьших квадратов (МНК), использующий следующие шаги для выбора подходящй модели:
+%\begin{enumerate}
+%\item для очередной модели:
+%\begin{itemize}
+%\item вычислить ошибки между рассматриваемой моделью и точками множества данных,
+%\item возвести в квадрат каждую ошибку,
+%\item просуммировать все квадраты ошибок,
+%\end{itemize}
+%\item выбрать модель, в которой сумма квадратов ошибок минимальна.
+%\end{enumerate}
+
+Комментарий 2.
+Оценка качества построенной модели на шаге 2 в пункте 4 производится с помощью метрики $R^2$.
+Значение $R^2$ или коэффициент детерминации - это единица минус доля необъяснённой дисперсии в дисперсии зависимой переменной.
+Она позволяет оценить эффективность моделей машинного обучения на основе регрессии, то есть характеризует схожесть построенной модели регрессии с целевой функцией в зафиксированном сечении.
+Значение $R^2$, как правило, находится между 0 и 1.
+Результат, близкий к 1, показывает, что точность модели высока.
+
+Комментарий 3.
+Значаения  $K$, $deg$ и $T$ являются параметрами метода и устанавливаются в соотвествии с особенностями решаемой задачей.
+В алгоритме возможно использование степени полинома регрессии $deg = 1$ или $deg = 2$, что позволяет построить модель линейной или квадратичной регрессии соотвественно.
+Для контроля определения "функция достаточно схожа с локальной" используется параметр метода $T$.
 
@@
 
-Описание тестовых задач
+В первом наборе тестовых задач (далее GRIS-L) использовалось 100 20-мерных задач комбинированного типа: локальная подзадача размерности 18, глобальная подзадача размерности 2.
+В качестве локальной составляющей использовались 18-мерные задачи, предсталяющие собой комбинацию почти линейных одномерных функций:
+\begin{equation}\label{X2_problem}
+\widetilde{\varphi}_1(y_1, ... , y_{18}) = \sum_{i=1}^{18} (M_{ij} y_i + L_{ij} y_i^2),
+\end{equation}
+где $-2.2 \leq y_1, .. ., y_{18} \leq 1.8$, а параметры $0.5 \leq |M_{ij}| \leq 1$ и $0.01 \leq L_{ij} \leq 0.3$  являются независимыми случайно равномерно распределёнными величинами.
+В качестве глобальной составляющей задачи использовались многоэкстремальные функции Гришагина \cite{}.
 
-Описание результатов экспериментов
+Во втором наборе задач (далее GRIS-LQ) параметры $M_{ij}$ и $L_{ij}$ локальной подзадачи (\ref{X2_problem}) были устроены таким образом, чтобы в 60\% случаев переменная имела несущественную квадратичную часть, в 30\% вклад квадратичной составляющей был существенным (как минимум, соизмеримым со вкладом линейной составляющей), а оставшиеся 10\% переменных были чисто линейными.
+
+В третьей серии экспериментов (далее GKLS-L) функции Гришагина заменялись 4-мерными функциями Сергеева \cite{}, в качестве локальной составляющей использовалась локальные подзадачи серии GRIS-L, размер серии был уменьшен до 20 задач.
+
+В результате решения серии GRIS-L с заданной точностью $0.01$ и параметрами метода $K=100$, $deg=1$ и $T=0.6$ во всех задачах было проведено корректное разделение на глобальные и локальные переменные, а затем задачи были успешно решены при помощи многошаговой схемы за приемлемое время и количество итераций.
+Однако решение усложнённой серии GRIS-LQ с применением также линейной регрессии не даёт хорошего результата, в силу наличия переменных параболического вида, для которых процент схожести с линейной моделью крайне низок.В связи с этим для проведения эксперимента были подобранны другие параметрв запуска $K=100$, $deg=2$ (квадратичная регрессия) и $T=0.95$.
+В результате все параметры задач серии были правильно разбиты на глобальные и локальные переменные, а затем успешно дорешены за приемлемое время и количество итераций.
+
+%\begin{figure}[ht]
+%\includegraphics[width=1.0\linewidth]{regression_94.png}
+%\caption{Построение регрессионных моделей на 94-й задаче серии GRIS-L}
+%\label{fig}
+%\end{figure}
+
+В третьей серии экспериментов в силу особенностей построения функций Сергеева успешно решить всю серию задач за приемлемое время и количество итераций не удалось.
+Большинство переменных функций GKLS в различных сечениях сильно подобны линейным и квадратичным функциям, отчего приходится сильно занижать порог вхождения $T$, чтобы случайно не отсеить глобальную переменную, как следствие часть локальных параметров также выносятся в глобальные.
+Большое количество глобальных параметров закономерно способствует увеличению количества итераций и времени решения задачи.
+С другой стороны, если поставить порог вхождения в локальные функции достаточно высоким, то глобальных переменных в некоторых функциях метод в принципе не обнаружит.
+
+Численные результаты серий экспериментов приведены в Таблице \ref{tab1}.
+
+\begin{table}[ht]
+	\caption{Результаты решения серий тестовых задач}
+	\label{tab1}
+		\begin{tabular}{ l c c c c c c } \hline
+		 & Количество &  Среднее & Среднее & Среднее & Среднее время \\ 
+		 & решенных  & число & число & число  & решения \\ 
+		 & задач & итераций & испытаний & запусков & задачи, с\\
+		 &  & & & локального & \\
+		 & & & & метода & \\ \hline
+		GRIS-L & 100/100  & 847 & 1692 & 451861 & 1,38 \\
+		GRIS-LQ & 100/100 & 751 & 1501 & 448326 & 1,21 \\
+		GKLS-L & 18/20 & 9531 & 19061 & 3899417 & 16,45 \\
+		\hline
+		\end{tabular}
+\end{table}
 
@@
 
-The calculation was carried out ...
+The calculation was carried out ... processor Intel Core™ i7 10750H, 6 ядер, 2,6 GHz.
 

### 2022-06-23 2022/ICNAAM/ICNAAM.tex
corrections in Russian

--- 2022-06-22 2022/ICNAAM/ICNAAM.tex
@@
 Таким образом, актуальной является разработка алгоритмов минимизации существенно многомерных функций, в которых учитывается разный характер зависимости целевой функции от разных групп параметров. 
-В данной работе предложен конкретный механизм такого разделения, основанный на <указать>.
+В данной работе предложен конкретный механизм такого разделения, основанный на анализе регрессионной модели функции по каждой переменной в отдельности.
 В качестве примеров, подтверждающих работоспособность предложенной схемы, решены как тестовые, так и прикладные задачи.
@@
 При этом разделение  параметров является неизвестным.
-Такой характер зависимостей является типичным, например, для обратных задач химической кинетики [ссылка].
 В данной работе предлагается подход к разделению параметров задачи на указанные две группы, конкретный механизм описан в следуюем разделе.
@@
 
-Описать алгоритм разделения параметров и формирования подзадач для многошаговой схемы
+Для разделения параметров задачи на локальные и глобальные предлагается следующая последовательность действий.
+
+Этап 1. Зафиксировать опорную точку $\overline{y} = (\overline{y}_1, \overline{y}_2,...,\overline{y}_N)$ целевой функции $\varphi(y)$. Выбор точки осуществляется одним из следующих способов:
+\begin{enumerate}
+\item точка фиксируется исходя из физического смысла решаемой прикладной задачи;
+\item в качестве фиксированной точки используется решение, найденное локальным методом за $K$ итераций (нашем исследовании использовался Hooke-Jeevse pattern search method).
+\item в качестве фиксированной точки используется решение, найденное глобальным методом за $K$ итераций (нашем исследовании использовался global search algorithm).
+\end{enumerate}
+
+Этап 2. Для всех компонент $y_i, \; 1\leq i \leq N$, провести исследование на локальность:
+\begin{enumerate}
+\item вычислить $P+1$ значение целевой функции $\varphi(y)$ в точках $z_i^j = (\overline{y}_1,...,\overline{y}_{i-1},y_i^j,\overline{y}_{i+1},...,\overline{y}_N)$, где
+$y_i^j =  a_i + jh, \; h=(b_i-a_i)/P, \; 0\leq j \leq P$;
+\item сформировать множество $Q$, представляющее собой набор из $P+1$ пары вида $\left\{y_i^j, \varphi(z_i^j)\right\} $ 
+%\[
+%Q = \left\{ (\overline{y}^k_i,  \varphi({\overline{y}^k})): \overline{y}^k = (\overline{y}_1, \overline{y}_2, … , \overline{y}^k_i, … ,\overline{y}_N), \overline{y}^k_i \sim U[a_i, b_i], 1\leq i \leq N, 1\leq k \leq P \right\};
+%\]
+\item построить регрессионную модель заданной степени $deg$ на основе данных из $Q$;
+\item вычислить оценку $R^2$ и провести классификацию переменной:
+\begin{itemize}
+\item если $R^2$ больше заданного порога $T$, то добавить переменную к множеству локальных переменных,
+\item иначе добавить переменную к множеству глобальных переменных.
+\end{itemize}
+\end{enumerate}
+
+Этап 3.
+Если локальные переменные были обнаружены, то запустить многошаговую схему, на верхнем уровне которой глобальные переменные, а на нижнем -- локальные.
+В противном случае запустить обычный алгоритм глобального поиска, считая все переменные глобальными.
+
+В основе проводимой регрессии лежит метод наименьших квадратов (МНК).
+%
+%Комментарий 1. В основе проводимой регрессии лежит метод наименьших квадратов (МНК), использующий следующие шаги для выбора подходящй модели:
+%\begin{enumerate}
+%\item для очередной модели:
+%\begin{itemize}
+%\item вычислить ошибки между рассматриваемой моделью и точками множества данных,
+%\item возвести в квадрат каждую ошибку,
+%\item просуммировать все квадраты ошибок,
+%\end{itemize}
+%\item выбрать модель, в которой сумма квадратов ошибок минимальна.
+%\end{enumerate}
+
+Оценка качества построенной модели на шаге 2 в пункте 4 производится с помощью коэффициента детерминации $R^2$.
+Указанный коэффициент характеризует схожесть построенной модели регрессии с целевой функцией в зафиксированном сечении.
+Значение $R^2$, как правило, находится между 0 и 1.
+Результат, близкий к 1, показывает, что точность модели высока.
+
+Значаения  $K$, $deg$ и $T$ являются параметрами метода и устанавливаются исследователем в соотвествии со свойствами решаемой задачей.
+В алгоритме возможно использование степени полинома регрессии $deg = 1$ или $deg = 2$, что позволяет построить модель линейной или квадратичной регрессии соотвественно.
+%Для контроля определения "функция достаточно схожа с локальной" используется параметр метода $T$.
 
@@
 
-Описание тестовых задач
+В первом наборе тестовых задач (будем обозначать его GRIS-L) использовалось 100 20-мерных задач комбинированного типа: локальная подзадача размерности 18, глобальная подзадача размерности 2.
+В качестве локальной составляющей использовались 18-мерные подзадачи, предсталяющие собой комбинацию близких к линейным одномерных функций:
+\begin{equation}\label{X2_problem}
+\overline{\varphi}_j(y) = \sum_{i=1}^{18} \left(M_{ij} y_i + L_{ij} y_i^2\right) ,
+\end{equation}
+где $-2.2 \leq y_i \leq 1.8$, а параметры $0.5 \leq |M_{ij}| \leq 1$ и $0.01 \leq L_{ij} \leq 0.3$  распределены случайно равномерно в соответствующих интервалах.
+В качестве глобальной составляющей задачи использовались многоэкстремальные функции, предложенные Гришагиным (см., например, \cite{Grishagin1994}).
 
-Описание результатов экспериментов
+Во втором наборе задач (далее GRIS-LQ) параметры $M_{ij}$ и $L_{ij}$ локальной подзадачи (\ref{X2_problem}) были устроены таким образом, чтобы в 60\% случаев переменная имела несущественную квадратичную часть, в 30\% вклад квадратичной составляющей был существенным (как минимум, соизмеримым со вкладом линейной составляющей), а оставшиеся 10\% переменных были чисто линейными.
 
-%\begin{figure}%[ht]
-%\includegraphics[width=1.0\linewidth]{fig1.png}
-%\caption{Schematic representation of the computational domain, with the geometry parameterization points marked in red and the black points independent of the optimization iterations.
-The grey lines show the possible deformation of the geometry.}
+В третьей серии экспериментов (далее GKLS-L) функции Гришагина заменялись 4-мерными функциями из класса GKLS (см., например, в \cite{Sergeyev2015,Grishagin2018}), в качестве локальной составляющей использовалась локальные подзадачи серии GRIS-L, размер серии составлял 20 задач.
+
+В результате решения серии GRIS-L с заданной точностью $0.01$ и параметрами метода $K=100$, $deg=1$ и $T=0.6$ во всех задачах было проведено корректное разделение на глобальные и локальные переменные, а затем задачи были успешно решены при помощи многошаговой схемы за приемлемое время и количество итераций.
+Однако решение усложнённой серии GRIS-LQ с применением также линейной регрессии не даёт хорошего результата, в силу наличия переменных параболического вида, для которых процент схожести с линейной моделью крайне низок.
+В связи с этим для проведения эксперимента были подобранны другие параметрв запуска $K=100$, $deg=2$ (квадратичная регрессия) и $T=0.95$.
+В результате все параметры задач серии были правильно разбиты на глобальные и локальные переменные, а затем успешно дорешены за приемлемое время и количество итераций.
+
+%\begin{figure}[ht]
+%\includegraphics[width=1.0\linewidth]{regression_94.png}
+%\caption{Построение регрессионных моделей на 94-й задаче серии GRIS-L}
 %\label{fig}
@@
 
-The calculation was carried out ...
+В третьей серии экспериментов в силу особенностей построения функций GKLS успешно решить всю серию задач за приемлемое время и количество итераций не удалось.
+Большинство переменных функций GKLS в различных сечениях сильно подобны линейным и квадратичным функциям, отчего приходится завышать пороговое значение $T$, чтобы случайно не отсеять глобальную переменную, как следствие часть локальных параметров также выносятся в глобальные.
+Большое количество глобальных параметров закономерно способствует увеличению количества итераций и времени решения задачи.
+С другой стороны, если поставить порог вхождения в локальные функции достаточно низким, то глобальных переменных в некоторых функциях метод в принципе не обнаружит.
+
+Численные результаты серий экспериментов приведены в Таблице \ref{tab1}.
+В ней представлены данные о числе решенных задач $S$, среднем числе итераций глобального поиска на верхнем уровне рекурсии $I_{av}$, среднем числе испытаний на нижнем уровне $T_{av}$, среднем времени решения задачи $t_{av}$.
+Вычисления проводились на компьютере с CPU Intel Core™ i7 10750H, 2.6 GHz.
+Для построения и анализа регрессионной модели использвался пакет scikit-learn из Python; метод глобальной оптмизации и nested optimization scheme реализованы на C++.
+
+\begin{table}[ht]
+	\caption{Результаты решения серий тестовых задач}
+	\label{tab1}
+		\begin{tabular}{ l c c c c c c } \hline
+		 & $S$ &  $I_{av}$ &  $T_{av}$ & $t_{av}, sec.$ \\
+    \hline
+		GRIS-L & 100/100  & 847 &  451 861 & 1.4 \\
+		GRIS-LQ & 100/100 & 751 &  448 326 & 1.2 \\
+		GKLS-L & 18/20 & 9531 &  3 899 417 & 16.4 \\
+		\hline
+		\end{tabular}
+\end{table}
 

### 2022-06-24 2022/ICNAAM/ICNAAM.tex
all text in Russian

@@
 При этом на разных уровнях рекурсии используются разные оптимизационные алгоритмы.
-Предложенный подход показал свою работоспособность при решении тестовых и прикладных задач.
+Предложенный подход показал свою работоспособность при решении нескольких серий тестовых задач.
 
@@
 В проведенном исследовании для решения задач глобальной оптимизации (\ref{global_problem}) мы использовали global search algorithm в сочетании со схемой редукции размерности на основе кривых Пеано \cite{Sergeyev2013,Lera2021}.
-Данный алгорим зарекомендовал себя как мощный инструмент для решения сложных многоэкстремальных задач \cite{Kvasov2013,Kalyulin2017,Cavoretto2021}.
+Данный алгоритм зарекомендовал себя как мощный инструмент для решения сложных многоэкстремальных задач \cite{Kvasov2013,Kalyulin2017,Cavoretto2021}.
 Для решения локальной задачи (\ref{local_problem}) мы использовали Hooke-Jeevse pattern search method (see, e.g., \cite{Kelley}).
@@
 
-Этап 1. Зафиксировать опорную точку $\overline{y} = (\overline{y}_1, \overline{y}_2,...,\overline{y}_N)$ целевой функции $\varphi(y)$. Выбор точки осуществляется одним из следующих способов:
+Этап 1. Зафиксировать опорную точку $\overline{y} = (\overline{y}_1, \overline{y}_2,...,\overline{y}_N)$ для целевой функции $\varphi(y)$. Выбор точки осуществляется одним из следующих способов:
 \begin{enumerate}
@@
 \item построить регрессионную модель заданной степени $deg$ на основе данных из $Q$;
-\item вычислить оценку $R^2$ и провести классификацию переменной:
+\item вычислить оценку $R^2$ и провести классификацию i-й переменной:
 \begin{itemize}
@@
 Этап 3.
-Если локальные переменные были обнаружены, то запустить многошаговую схему, на верхнем уровне которой глобальные переменные, а на нижнем -- локальные.
-В противном случае запустить обычный алгоритм глобального поиска, считая все переменные глобальными.
+Если локальные переменные были обнаружены, то продолжить решение задачи с использованием nested optimization scheme, на верхнем уровне которой будет использоваться алгоритм глобального поиска, а на нижнем уровне -- метод локальной оптимизации.
+В противном случае запустить алгоритм глобального поиска, считая все переменные глобальными.
 
-В основе проводимой регрессии лежит метод наименьших квадратов (МНК).
 %
@@
 
-Оценка качества построенной модели на шаге 2 в пункте 4 производится с помощью коэффициента детерминации $R^2$.
+В основе проводимой регрессии лежит метод наименьших квадратов (МНК).
+Оценка качества построенной модели на шаге 2 в пункте 4 производится с помощью adjusted coefficient of determination $R^2_a$.
 Указанный коэффициент характеризует схожесть построенной модели регрессии с целевой функцией в зафиксированном сечении.
-Значение $R^2$, как правило, находится между 0 и 1.
+Значение $R^2_a$, как правило, находится между 0 и 1.
 Результат, близкий к 1, показывает, что точность модели высока.
@@
 Значаения  $K$, $deg$ и $T$ являются параметрами метода и устанавливаются исследователем в соотвествии со свойствами решаемой задачей.
-В алгоритме возможно использование степени полинома регрессии $deg = 1$ или $deg = 2$, что позволяет построить модель линейной или квадратичной регрессии соотвественно.
+В алгоритме возможно использование степени полинома регрессии $deg = 1$ или $deg = 2$, что позволяет построить модель линейной или квадратичной регрессии соответственно.
 %Для контроля определения "функция достаточно схожа с локальной" используется параметр метода $T$.
@@
 В первом наборе тестовых задач (будем обозначать его GRIS-L) использовалось 100 20-мерных задач комбинированного типа: локальная подзадача размерности 18, глобальная подзадача размерности 2.
-В качестве локальной составляющей использовались 18-мерные подзадачи, предсталяющие собой комбинацию близких к линейным одномерных функций:
+В качестве локальной составляющей использовались 18-мерные подзадачи, представляющие собой комбинацию близких к линейным одномерных функций:
 \begin{equation}\label{X2_problem}
@@
 В качестве глобальной составляющей задачи использовались многоэкстремальные функции, предложенные Гришагиным (см., например, \cite{Grishagin1994}).
+%Марина, где-то здесь надо сказать, как формировалась целевая функция - общий вид, "перемешанные" параметры, и т.п. Буквально 1-2 предложения, не больше
 
@@
 
-В третьей серии экспериментов (далее GKLS-L) функции Гришагина заменялись 4-мерными функциями из класса GKLS (см., например, в \cite{Sergeyev2015,Grishagin2018}), в качестве локальной составляющей использовалась локальные подзадачи серии GRIS-L, размер серии составлял 20 задач.
+В третьей серии экспериментов (далее GKLS-L) в качестве многоэкстремальной части использовались 4-мерные функции из класса GKLS (см., например, в \cite{Sergeyev2015,Grishagin2018}), в качестве локальной составляющей использовалась локальные подзадачи серии GRIS-L, размер серии составлял 20 задач.
 
-В результате решения серии GRIS-L с заданной точностью $0.01$ и параметрами метода $K=100$, $deg=1$ и $T=0.6$ во всех задачах было проведено корректное разделение на глобальные и локальные переменные, а затем задачи были успешно решены при помощи многошаговой схемы за приемлемое время и количество итераций.
-Однако решение усложнённой серии GRIS-LQ с применением также линейной регрессии не даёт хорошего результата, в силу наличия переменных параболического вида, для которых процент схожести с линейной моделью крайне низок.
-В связи с этим для проведения эксперимента были подобранны другие параметрв запуска $K=100$, $deg=2$ (квадратичная регрессия) и $T=0.95$.
-В результате все параметры задач серии были правильно разбиты на глобальные и локальные переменные, а затем успешно дорешены за приемлемое время и количество итераций.
+В результате решения серии GRIS-L с заданной точностью $0.01$ и параметрами метода $K=100$, $deg=1$ и $T=0.6$ во всех задачах было проведено корректное разделение на глобальные и локальные переменные, а затем задачи были успешно решены при помощи nested optimization scheme за приемлемое время и количество итераций.
+Однако решение усложнённой серии GRIS-LQ с применением линейной регрессии не даёт хорошего результата, в силу наличия переменных параболического вида, для которых процент схожести с линейной моделью крайне низок.
+В связи с этим для проведения эксперимента были подобраны другие параметрs запуска $K=100$, $deg=2$ (квадратичная регрессия) и $T=0.95$.
+В результате все параметры задач были правильно разбиты на глобальные и локальные, а сами задачи затем успешно решены за приемлемое время и количество итераций.
 
@@
 Численные результаты серий экспериментов приведены в Таблице \ref{tab1}.
-В ней представлены данные о числе решенных задач $S$, среднем числе итераций глобального поиска на верхнем уровне рекурсии $I_{av}$, среднем числе испытаний на нижнем уровне $T_{av}$, среднем времени решения задачи $t_{av}$.
+В ней представлены данные о числе решенных задач $S$, среднем числе итераций глобального поиска на верхнем уровне рекурсии $G_{av}$, среднем числе испытаний на нижнем уровне $L_{av}$, среднем времени решения задачи $t_{av}$.
 Вычисления проводились на компьютере с CPU Intel Core™ i7 10750H, 2.6 GHz.
-Для построения и анализа регрессионной модели использвался пакет scikit-learn из Python; метод глобальной оптмизации и nested optimization scheme реализованы на C++.
+Для построения и анализа регрессионной модели использовался пакет scikit-learn из Python; global search algorithm and nested optimization scheme реализованы на C++.
 
@@
 		\begin{tabular}{ l c c c c c c } \hline
-		 & $S$ &  $I_{av}$ &  $T_{av}$ & $t_{av}, sec.$ \\
+		 & $S$ &  $G_{av}$ &  $L_{av}$ & $t_{av}, sec.$ \\
     \hline

### 2022-06-24 2022/ICNAAM/ICNAAM.tex
add some sentence about objective function
before translation

@@
 В данной работе предложен конкретный механизм такого разделения, основанный на анализе регрессионной модели функции по каждой переменной в отдельности.
-В качестве примеров, подтверждающих работоспособность предложенной схемы, решены как тестовые, так и прикладные задачи.
+Для подтверждения работоспособность предложенной схемы решен представительный набор тестовых задач.
 
@@
 with the constant $L$ unknown a priori.
-Одновременно с этим будет предполагать, что зависимость целевой функции от $N_1$ параметров является многоэкстремальной, а от оставшихся $N_2 = N - N_1$ параметров зависимость являеся близкой к линейной.
+Одновременно с этим будет предполагать, что зависимость целевой функции от $N_1$ параметров является многоэкстремальной, а от оставшихся $N_2 = N - N_1$ параметров зависимость является близкой к линейной или квадратичной.
 При этом разделение  параметров является неизвестным.
-В данной работе предлагается подход к разделению параметров задачи на указанные две группы, конкретный механизм описан в следуюем разделе.
+В данной работе предлагается подход к разделению параметров задачи на указанные две группы, конкретный механизм описан в следующем разделе.
 
@@
 Эффективность и проработанность методов локальной оптимизации позволяет решать подзадачи (\ref{local_problem}) с точностью, значительно превышающей точность методов глобальной оптимизации.
-Соответственно, можно рассматривать задачу (\ref{global_problem}) как задачу глобальной оптимизации, в которой значение функции может быть вычислено с высокой точностью, но данная операция является трудоемкой.
+Соответственно, можно рассматривать задачу (\ref{global_problem}) как задачу глобальной оптимизации, в которой процесс вычисления значения целевой функции является трудоемкой операцией.
 
-В проведенном исследовании для решения задач глобальной оптимизации (\ref{global_problem}) мы использовали global search algorithm в сочетании со схемой редукции размерности на основе кривых Пеано \cite{Sergeyev2013,Lera2021}.
+В проведенном исследовании для решения задач глобальной оптимизации (\ref{global_problem}) мы использовали global search algorithm в сочетании со схемой редукции размерности на основе Peano curves \cite{Sergeyev2013,Lera2021}.
 Данный алгоритм зарекомендовал себя как мощный инструмент для решения сложных многоэкстремальных задач \cite{Kvasov2013,Kalyulin2017,Cavoretto2021}.
@@
 \item сформировать множество $Q$, представляющее собой набор из $P+1$ пары вида $\left\{y_i^j, \varphi(z_i^j)\right\} $ 
-%\[
-%Q = \left\{ (\overline{y}^k_i,  \varphi({\overline{y}^k})): \overline{y}^k = (\overline{y}_1, \overline{y}_2, … , \overline{y}^k_i, … ,\overline{y}_N), \overline{y}^k_i \sim U[a_i, b_i], 1\leq i \leq N, 1\leq k \leq P \right\};
-%\]
 \item построить регрессионную модель заданной степени $deg$ на основе данных из $Q$;
@@
 
-%
-%Комментарий 1. В основе проводимой регрессии лежит метод наименьших квадратов (МНК), использующий следующие шаги для выбора подходящй модели:
-%\begin{enumerate}
-%\item для очередной модели:
-%\begin{itemize}
-%\item вычислить ошибки между рассматриваемой моделью и точками множества данных,
-%\item возвести в квадрат каждую ошибку,
-%\item просуммировать все квадраты ошибок,
-%\end{itemize}
-%\item выбрать модель, в которой сумма квадратов ошибок минимальна.
-%\end{enumerate}
-
 В основе проводимой регрессии лежит метод наименьших квадратов (МНК).
@@
 
-Значаения  $K$, $deg$ и $T$ являются параметрами метода и устанавливаются исследователем в соотвествии со свойствами решаемой задачей.
+Значения  $K$, $deg$ и $T$ являются параметрами метода и устанавливаются исследователем в соответствии со свойствами решаемой задачей.
 В алгоритме возможно использование степени полинома регрессии $deg = 1$ или $deg = 2$, что позволяет построить модель линейной или квадратичной регрессии соответственно.
-%Для контроля определения "функция достаточно схожа с локальной" используется параметр метода $T$.
 
@@
 \begin{equation}\label{X2_problem}
-\overline{\varphi}_j(y) = \sum_{i=1}^{18} \left(M_{ij} y_i + L_{ij} y_i^2\right) ,
+\varphi_j^{loc}(y) = \sum_{i=1}^{18} \left(M_{ij} y_i + L_{ij} y_i^2\right) ,
 \end{equation}
 где $-2.2 \leq y_i \leq 1.8$, а параметры $0.5 \leq |M_{ij}| \leq 1$ и $0.01 \leq L_{ij} \leq 0.3$  распределены случайно равномерно в соответствующих интервалах.
-В качестве глобальной составляющей задачи использовались многоэкстремальные функции, предложенные Гришагиным (см., например, \cite{Grishagin1994}).
-%Марина, где-то здесь надо сказать, как формировалась целевая функция - общий вид, "перемешанные" параметры, и т.п. Буквально 1-2 предложения, не больше
+В качестве глобальной составляющей задачи $\varphi_j^{glob}(y)$ использовались многоэкстремальные функции, предложенные Гришагиным (см., например, \cite{Grishagin1994}).
+Целевая функция $\varphi(y)$  представляла собой линейную комбинацию указанных подзадач, т.е.
+$\varphi_j(y) = \alpha \varphi_j^{glob}(y) + \beta \varphi_j^{loc}(y)$ .
+Последовательность параметров в каждой задаче задавалась случайным образом.
 
@@
 Однако решение усложнённой серии GRIS-LQ с применением линейной регрессии не даёт хорошего результата, в силу наличия переменных параболического вида, для которых процент схожести с линейной моделью крайне низок.
-В связи с этим для проведения эксперимента были подобраны другие параметрs запуска $K=100$, $deg=2$ (квадратичная регрессия) и $T=0.95$.
+В связи с этим для проведения эксперимента были подобраны другие параметры запуска $K=100$, $deg=2$ (квадратичная регрессия) и $T=0.95$.
 В результате все параметры задач были правильно разбиты на глобальные и локальные, а сами задачи затем успешно решены за приемлемое время и количество итераций.
-
-%\begin{figure}[ht]
-%\includegraphics[width=1.0\linewidth]{regression_94.png}
-%\caption{Построение регрессионных моделей на 94-й задаче серии GRIS-L}
-%\label{fig}
-%\end{figure}
 
@@
 Численные результаты серий экспериментов приведены в Таблице \ref{tab1}.
-В ней представлены данные о числе решенных задач $S$, среднем числе итераций глобального поиска на верхнем уровне рекурсии $G_{av}$, среднем числе испытаний на нижнем уровне $L_{av}$, среднем времени решения задачи $t_{av}$.
+В ней представлены данные о числе решенных задач $S$, среднем числе итераций глобального поиска на верхнем уровне рекурсии $G_{av}$, среднем числе испытаний на нижнем уровне $L_{av}$, среднем времени решения задачи $t_{av}$ (в секундах).
 Вычисления проводились на компьютере с CPU Intel Core™ i7 10750H, 2.6 GHz.
@@
 		\begin{tabular}{ l c c c c c c } \hline
-		 & $S$ &  $G_{av}$ &  $L_{av}$ & $t_{av}, sec.$ \\
+		 & $S$ &  $G_{av}$ &  $L_{av}$ & $t_{av}$ \\
     \hline

### 2022-06-27 2022/ICNAAM/ICNAAM.tex
after translation

@@
 
-%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
-\usepackage{inputenc}
-\usepackage[T2A,T1]{fontenc}
-\usepackage[english,russian]{babel}
-\usepackage{cmap}
+%% ƒÀﬂ –”—— Œ√Œ “≈ —“¿ Á‡ÍÓÏÏÂÌÚËÓ‚‡Ú¸ ÔÓÚÓÏ!
+%\usepackage{inputenc}
+%\usepackage[T2A,T1]{fontenc}
+%\usepackage[english,russian]{babel}
+%\usepackage{cmap}
 %%
@@
 % Title portion
-\title{Использование методов машинного обучения для разделения параметров в задачах глобальной оптимизации большой размерности}
+\title{Using Machine Learning for Subdivision of Parameters in Global Optimization Problems of Large Dimensionality}
 
@@
 \begin{abstract}
-В работе представлены результаты исследования подхода к решению задач оптимизации, в которых зависимость от разных групп параметров носит разный характер.
-Предложена схема выделения параметров задачи, которые оказывают локальное влияние на целевую функцию, что позволяет решать существенно многомерные задачи с использованием nested optimization scheme.
-При этом на разных уровнях рекурсии используются разные оптимизационные алгоритмы.
-Предложенный подход показал свою работоспособность при решении нескольких серий тестовых задач.
+The work presents the results of investigations of an approach to solving the optimization problems, in which the dependencies on different groups of parameters have different nature.
+A scheme for selecting the parameters of the problem, which have local effect on the objective function, has been proposed.
+It allows to solve the essentially multidimensional problems using the nested optimization scheme.
+At that, different optimization algorithms are used at different nesting levels.
+The proposed approach has demonstrated its efficiency in solving several series of test problems.
 
@@
 
-В настоящее время методы глобальной оптимизации используются для решения широкого круга задач, возникающих в различных областях науки и техники. 
-Например, традиционной сферой применения методов глобальной оптимизации является идентификация значений параметров математических моделей по данным экспериментов.
-К их числу относятся, например, inverse problems of chemical kinetics \cite{Gubaydullin2021}.
-В задачах такого вида требуется провести поиск значений неизвестных параметров модели, при которых результаты моделирования наиболее близки к результатам, полученным экспериментально.
+At present, the global optimization methods are applied to solving various problems arising in various fields of science and technology. 
+For example, the evaluation of the model parameters from the experimental data is a traditional field of application of the global optimization methods.
+Among these ones, there are, for example, the problems of chemical kinetics \cite{Gubaydullin2021}.
+In the problems of this kind, it is necessary to perform a search for the values of unknown model parameters, at which the modeling results are the closest to 
+the ones obtained experimentally.
 
-Число параметров, которые требуется идентифицировать подобным образом, для задач химической кинетики может составлять десятки и сотни.
-Использование детерминированных методов глобальной оптимизации для решения задачах такой размерности крайне ограничено из-за чрезвычайно больших вычислительных затрат на покрытие search domain точками испытаний.
-Это остается справедливым даже в случае использования эффективных алгоритмов (например, \cite{Paulavicius2011,Evtushenko2009,Jones2009}), строящих существенно неравномерные покрытия.
+The number of parameters, which are to be evaluated in this way, can be tens and hundreds for the chemical kinetics problems.
+The use of the deterministic global optimization methods for solving the problems of such a dimensionality is very limited because of very large computation costs of covering the search domain by the trial points.
+This holds even in the case of using the efficient algorithms (for example, \cite{Paulavicius2011,Evtushenko2009,Jones2009}) constructing the essentially nonuniform coverages.
 
-Многие обратные задачи характеризуются тем, что зависимость от разных групп параметров носит разный характер.
-Например, от одной группы параметров зависимость может быть близка к линейной, тогда как по второй группе зависимость может носить сложный многоэкстремальный характер.
+Numerous inverse problems are featured by different dependencies on different groups of parameters. 
+For example, the dependence on one group of parameters may be close to the linear one whereas the dependence on another group of parameters may have a complex multiextremal character.
 
-В этом случае решение задачи можно организовать по схеме recursive optimization.
-Решение  многоэкстремальной подзадачи, для которой требуется использовать сложные алгоритмы глобальной оптимизации, будет проводиться на верхнем уровне рекурсии.
-Унимодальные подзадачи (каждая из которых соответствует фиксированному набору значений первой части параметров) будут решаться на нижнем уровне.
-Для их решения можно применять обладающие быстрой сходимостью методы локальной оптимизации.
+In this case, the problem can be solved using the recursive optimization scheme. 
+The solving of the multiextremal subproblem, which requires the use of complex global optimization algorithms, will be performed at the upper nesting level.
+The unimodal subproblems (each corresponding to a fixed set of the values of the first part of parameters) will be solve at the lower nesting level.
+To solve these ones, one can apply the local optimization methods having a fast convergence.
 
-Однако заранее указать разделение на группы параметров с разным характером поведения целевой функции, как правило, не представляется возможным, т.к.
-целевая функция в обратных задачах задается как черный ящик.
-Таким образом, актуальной является разработка алгоритмов минимизации существенно многомерных функций, в которых учитывается разный характер зависимости целевой функции от разных групп параметров. 
-В данной работе предложен конкретный механизм такого разделения, основанный на анализе регрессионной модели функции по каждой переменной в отдельности.
-Для подтверждения работоспособность предложенной схемы решен представительный набор тестовых задач.
+As a rule, one can't subdivide the parameters into the groups with different character of the objective function behavior in advance since the objective function in applied problems is usually defined as a black box.
+So far, the development of the algorithm for the minimization of the essentially multiextremal functions taking into account different character of the dependence of the objective function on different groups of parameters is relevant.
+In the present work, a special method for such a subdivision based on the analysis of the regression model of the function with respect to each variable separately has been proposed.
+To confirm the efficiency of the scheme proposed, a representative set of test problems has been solved.
 
-\section{ОПТИМИЗАЦИОННЫЕ АЛГОРИТМЫ И РЕДУКИЦЯ РАЗМЕРНОСТИ}
+\section{OPTIMIZATION ALGORITHMS AND DIMENSIONALITY REDUCTION}
 
-Рассмотрим задачу оптимизации вида
+The paper considers an optimization problem of the form 
 \begin{eqnarray}\label{main_problem}
@@
 \end{eqnarray}
-Будем предполагать, что функция $\varphi(y)$ является многоэкстремальной и удовлетворяет Lipschitz condition
+
+Let us assume the function $\varphi(y)$ to be a multiextremal one and to satisfy the Lipschitz condition
 \[
@@
 \]
-with the constant $L$ unknown a priori.
-Одновременно с этим будет предполагать, что зависимость целевой функции от $N_1$ параметров является многоэкстремальной, а от оставшихся $N_2 = N - N_1$ параметров зависимость является близкой к линейной или квадратичной.
-При этом разделение  параметров является неизвестным.
-В данной работе предлагается подход к разделению параметров задачи на указанные две группы, конкретный механизм описан в следующем разделе.
+with the constant $L$ unknown {\em a priori}.
+At the same time, let us assume the dependence of the objective function on $N_1$ parameters to be multiextremal while the dependence on the rest $N_2 = N - N_1$ parameters is close to the linear or quadratic one.
+At the same time, the subdivision of the parameters is unknown. 
+In the present work, an approach to the subdivision of the problem parameters into the two specified groups is proposed.
+The details of the method are described in the next section.
 
-Учет подобной особенности задач может существенно снизить вычислительную сложность поиска оптимума.
-В самом деле, в соответствии с известной схемой рекурсивной оптимизации \cite{Carr} решение исходной задачи можно проводить следующим образом:
+Taking into account such a specific feature of the problem can essentially reduce the computation costs of the search for the optimum. 
+Indeed, according to well known nested optimization scheme \cite{Grishagin2001,Grishagin2015}, the solving of the initial problem can be performed in the following way:
 \begin{equation}\label{global_problem}
-\varphi(y^*) = \min_{y\in D} \varphi (y) = \min_{y^1\in D_1} f(y^1), \; D_1=\left\{y\in R^{N_1}: a_i\leq y_i \leq b_i, \; 1\leq i \leq N_1\right\},
+\varphi(y^*) = \min_{y\in D} \varphi (y) = \min_{y^1\in D_1} f(y^1), \; D_1=\left\{y\in R^{N_1}: 
+a_i\leq y_i \leq b_i, \; 1\leq i \leq N_1\right\},
 \end{equation}
-где 
+where 
 \begin{equation}\label{local_problem}
-f(y^1) = \min_{ y^2 \in D_2} \varphi(y^1,y^2), \; D_2=\left\{y\in R^{N_2}: a_i\leq y_i \leq b_i, \; N_1+1\leq i \leq N\right\}.
+f(y^1) = \min_{ y^2 \in D_2} \varphi(y^1,y^2), \; D_2=\left\{y\in R^{N_2}: a_i\leq y_i \leq b_i, \; 
+N_1+1\leq i \leq N\right\}.
 \end{equation}
-В соответствии с (\ref{global_problem}) вычисление одного значения функции $\varphi^1 (y^1)$ (данный процесс будем называть search trial) подразумевает решение унимодальной задачи (\ref{local_problem}), которое может быть выполнено одним из методов локальной оптимизации.
-Эффективность и проработанность методов локальной оптимизации позволяет решать подзадачи (\ref{local_problem}) с точностью, значительно превышающей точность методов глобальной оптимизации.
-Соответственно, можно рассматривать задачу (\ref{global_problem}) как задачу глобальной оптимизации, в которой процесс вычисления значения целевой функции является трудоемкой операцией.
 
-В проведенном исследовании для решения задач глобальной оптимизации (\ref{global_problem}) мы использовали global search algorithm в сочетании со схемой редукции размерности на основе Peano curves \cite{Sergeyev2013,Lera2021}.
-Данный алгоритм зарекомендовал себя как мощный инструмент для решения сложных многоэкстремальных задач \cite{Kvasov2013,Kalyulin2017,Cavoretto2021}.
-Для решения локальной задачи (\ref{local_problem}) мы использовали Hooke-Jeevse pattern search method (see, e.g., \cite{Kelley}).
-В отличие от методов типа BFGS (see, e.g., \cite{Nocedal}), данный алгоритм не требует вычисления градиента и хорошо работает в условиях наличия вычислительных ошибок в значениях функций, что является типичным при решении прикладных задач.
+According to (\ref{global_problem}), the computation of a single value of the function $f(y^1)$
+%(this process will be called {\em a search trial})
+implies the solving of a unimodal problem (\ref{local_problem}), which can be performed by one of local optimization methods.
+The efficiency of the local optimization methods allow solving the subproblems (\ref{local_problem}) with a precision essentially exceeding the one of the global optimization methods.
+Accordingly, we can consider the problem (\ref{global_problem}) as the global optimization one where the process of computing an objective function value to be a 
+time-consuming operation.
 
-\section{РАЗДЕЛЕНИЕ ПАРАМЕТРОВ ЗАДАЧИ}
+In the present study, we used the global search algorithm combined with the dimensionality reduction scheme based on Peano curves \cite{Sergeyev2013,Lera2021} to solve the global optimization problem (\ref{global_problem}).
+This algorithm was demonstrated to be a powerful tool for solving complex multiextremal problems \cite{Kvasov2013,Kalyulin2017,Cavoretto2021}.
+To solve the local problem (\ref{local_problem}), we used Hooke-Jeevse pattern search method (see, e.g., \cite{Kelley}). 
+Unlike the BFGS-type methods (see, e.g., \cite{Nocedal}), this algorithm doesn't require computing the gradient and works well under conditions of the 
+computing errors in the function values that is typical in solving the applied problems.
 
-Для разделения параметров задачи на локальные и глобальные предлагается следующая последовательность действий.
+\section{SUBDIVISION OF THE PROBLEMS PARAMETERS}
 
-Этап 1. Зафиксировать опорную точку $\overline{y} = (\overline{y}_1, \overline{y}_2,...,\overline{y}_N)$ для целевой функции $\varphi(y)$. Выбор точки осуществляется одним из следующих способов:
+To divide the problem parameters into the local and global ones, we propose the following sequence of operations.
+
+Step 1.
+Fix the reference point $\overline{y} = (\overline{y}_1, \overline{y}_2,...,\overline{y}_N)$ for the objective function $\varphi(y)$.
+The choice of the point is performed by one of the following methods:
 \begin{enumerate}
-\item точка фиксируется исходя из физического смысла решаемой прикладной задачи;
-\item в качестве фиксированной точки используется решение, найденное локальным методом за $K$ итераций (нашем исследовании использовался Hooke-Jeevse pattern search method).
-\item в качестве фиксированной точки используется решение, найденное глобальным методом за $K$ итераций (нашем исследовании использовался global search algorithm).
+\item the point is fixed according to the physical meaning of the applied problem being solved;
+\item use the solution found by the local method in $K$ iterations (in the present study, Hooke-Jeevse pattern search method was used) as the reference point.
+\item use the solution found by the global method in $K$ iterations (in the present study, the global search algorithm was used) as the reference point.
 \end{enumerate}
 
-Этап 2. Для всех компонент $y_i, \; 1\leq i \leq N$, провести исследование на локальность:
+Step 2. For all components of $y_i, \; 1\leq i \leq N$, investigate the locality:
 \begin{enumerate}
-\item вычислить $P+1$ значение целевой функции $\varphi(y)$ в точках $z_i^j = (\overline{y}_1,...,\overline{y}_{i-1},y_i^j,\overline{y}_{i+1},...,\overline{y}_N)$, где
-$y_i^j =  a_i + jh, \; h=(b_i-a_i)/P, \; 0\leq j \leq P$;
-\item сформировать множество $Q$, представляющее собой набор из $P+1$ пары вида $\left\{y_i^j, \varphi(z_i^j)\right\} $ 
-\item построить регрессионную модель заданной степени $deg$ на основе данных из $Q$;
-\item вычислить оценку $R^2$ и провести классификацию i-й переменной:
+\item compute $P+1$ values of the objective function $\varphi(y)$ at the points $z_i^j = (\overline{y}_1,...,\overline{y}_{i-1},y_i^j,\overline{y}_{i+1},...,\overline{y}_N)$ where $y_i^j =  a_i + jh, \; h=(b_i-a_i)/P, \; 0\leq j \leq P$;
+\item form a set $Q$ comprising of $P+1$ pairs of the kind $\left\{y_i^j, \varphi(z_i^j)\right\} $ 
+\item construct a regression model of given power $deg$ based on the data from $Q$;
+\item compute the coefficient of determination $R^2$ and perform the classification of the $i^{\rm th}$ variable:
 \begin{itemize}
-\item если $R^2$ больше заданного порога $T$, то добавить переменную к множеству локальных переменных,
-\item иначе добавить переменную к множеству глобальных переменных.
+\item if $R^2$ is greater than a predefined threshold $T$, then add the variable to the set of the local variables,
+\item else add the variable to the set of the global variables.
 \end{itemize}
@@
 
-Этап 3.
-Если локальные переменные были обнаружены, то продолжить решение задачи с использованием nested optimization scheme, на верхнем уровне которой будет использоваться алгоритм глобального поиска, а на нижнем уровне -- метод локальной оптимизации.
-В противном случае запустить алгоритм глобального поиска, считая все переменные глобальными.
+Step 3.
+If the local variables are found, then continue solving the problem using the nested optimization scheme; the global search algorithm will be used at the upper level while the local optimization method will be used at the lower one.
+Otherwise, run the global search algorithm taking all variables as the global ones.
 
-В основе проводимой регрессии лежит метод наименьших квадратов (МНК).
-Оценка качества построенной модели на шаге 2 в пункте 4 производится с помощью adjusted coefficient of determination $R^2_a$.
-Указанный коэффициент характеризует схожесть построенной модели регрессии с целевой функцией в зафиксированном сечении.
-Значение $R^2_a$, как правило, находится между 0 и 1.
-Результат, близкий к 1, показывает, что точность модели высока.
+The regression is based on the least squares method (LSM).
+The quality of the model constructed was evaluated at Step 2, Item 4 using the adjusted coefficient of determination $R^2_a$.
+This coefficient features the similarity of the regression model constructed to the objective function in the fixed section.
+The value of $R^2_a$, as a rule, falls between 0 and 1.
+A result close to 1 shows the accurasy of the model to be high.
 
-Значения  $K$, $deg$ и $T$ являются параметрами метода и устанавливаются исследователем в соответствии со свойствами решаемой задачей.
-В алгоритме возможно использование степени полинома регрессии $deg = 1$ или $deg = 2$, что позволяет построить модель линейной или квадратичной регрессии соответственно.
+The values of $K$, $deg$, and $T$ are the parameters of the method and are set by the researcher according to the properties of the problem being solved.
+The algorithm allows using the power of the regression polynomial $deg = 1$ or $deg = 2$ that allows constructing the model based on either linear or quadratic regression, respectively.
 
@@
 
-В первом наборе тестовых задач (будем обозначать его GRIS-L) использовалось 100 20-мерных задач комбинированного типа: локальная подзадача размерности 18, глобальная подзадача размерности 2.
-В качестве локальной составляющей использовались 18-мерные подзадачи, представляющие собой комбинацию близких к линейным одномерных функций:
+In the first set of the test problems (we will denote it as Gr-L), we used 100 20-dimensional combined-type problems: the local subproblems of the dimensionality $N_2 = 18$ and a global subproblem of the dimensionality $N_1=2$.
+The 18-dimensional subproblems comprising the combinations of nearly linear functions were used as the local component:
 \begin{equation}\label{X2_problem}
-\varphi_j^{loc}(y) = \sum_{i=1}^{18} \left(M_{ij} y_i + L_{ij} y_i^2\right) ,
+\varphi_j^{loc}(y) = \sum_{i=1}^{18} \left(M_{ij} y_i + L_{ij} y_i^2\right), \; -2.2 \leq y_i \leq 1.8,
 \end{equation}
-где $-2.2 \leq y_i \leq 1.8$, а параметры $0.5 \leq |M_{ij}| \leq 1$ и $0.01 \leq L_{ij} \leq 0.3$  распределены случайно равномерно в соответствующих интервалах.
-В качестве глобальной составляющей задачи $\varphi_j^{glob}(y)$ использовались многоэкстремальные функции, предложенные Гришагиным (см., например, \cite{Grishagin1994}).
-Целевая функция $\varphi(y)$  представляла собой линейную комбинацию указанных подзадач, т.е.
-$\varphi_j(y) = \alpha \varphi_j^{glob}(y) + \beta \varphi_j^{loc}(y)$ .
-Последовательность параметров в каждой задаче задавалась случайным образом.
+where the parameters $0.5 \leq |M_{ij}| \leq 1$ and $0.01 \leq L_{ij} \leq 0.3$  were randomly uniformly distributed within the respective intervals.
+The multiextremal functions proposed by Grishagin (see, for example, \cite{Grishagin1994}) were used as the global components of the problems $\varphi_j^{glob}(y)$.
 
-Во втором наборе задач (далее GRIS-LQ) параметры $M_{ij}$ и $L_{ij}$ локальной подзадачи (\ref{X2_problem}) были устроены таким образом, чтобы в 60\% случаев переменная имела несущественную квадратичную часть, в 30\% вклад квадратичной составляющей был существенным (как минимум, соизмеримым со вкладом линейной составляющей), а оставшиеся 10\% переменных были чисто линейными.
+The objective function $\varphi(y)$ comprised a linear combination of these subproblems, i.e.
+$\varphi_j(y) = \alpha \varphi_j^{glob}(y) + \beta \varphi_j^{loc}(y)$.
+The sequence of the parameters in each problem was selected randomly.
 
-В третьей серии экспериментов (далее GKLS-L) в качестве многоэкстремальной части использовались 4-мерные функции из класса GKLS (см., например, в \cite{Sergeyev2015,Grishagin2018}), в качестве локальной составляющей использовалась локальные подзадачи серии GRIS-L, размер серии составлял 20 задач.
+In the second set of problems (hereafter called Gr-LQ), the parameters $M_{ij}$ and $L_{ij}$ of the local subproblem (\ref{X2_problem}) were organized so that in 60\% of the cases the viable had an insufficient quadratic part, in 30\% of the cases the contributions of the quadratic components were essential (at least, comparable to the contribution of the linear parts), and the rest 10\% of variables were purely linear.
 
-В результате решения серии GRIS-L с заданной точностью $0.01$ и параметрами метода $K=100$, $deg=1$ и $T=0.6$ во всех задачах было проведено корректное разделение на глобальные и локальные переменные, а затем задачи были успешно решены при помощи nested optimization scheme за приемлемое время и количество итераций.
-Однако решение усложнённой серии GRIS-LQ с применением линейной регрессии не даёт хорошего результата, в силу наличия переменных параболического вида, для которых процент схожести с линейной моделью крайне низок.
-В связи с этим для проведения эксперимента были подобраны другие параметры запуска $K=100$, $deg=2$ (квадратичная регрессия) и $T=0.95$.
-В результате все параметры задач были правильно разбиты на глобальные и локальные, а сами задачи затем успешно решены за приемлемое время и количество итераций.
+In the third series of experiments (hereafter called GKLS-L), 4-dimensional functions from the GKLS class (see, for example, in \cite{Sergeyev2015,Grishagin2018}) were used as the multiextremal parts.
+As the local components, the local subproblems of the Gr-L series were used.
+The series included 20 problems.
 
-В третьей серии экспериментов в силу особенностей построения функций GKLS успешно решить всю серию задач за приемлемое время и количество итераций не удалось.
-Большинство переменных функций GKLS в различных сечениях сильно подобны линейным и квадратичным функциям, отчего приходится завышать пороговое значение $T$, чтобы случайно не отсеять глобальную переменную, как следствие часть локальных параметров также выносятся в глобальные.
-Большое количество глобальных параметров закономерно способствует увеличению количества итераций и времени решения задачи.
-С другой стороны, если поставить порог вхождения в локальные функции достаточно низким, то глобальных переменных в некоторых функциях метод в принципе не обнаружит.
+As a result of solving the Gr-L series with the preset precision of $0.01$ and the parameters of the method $K=100$, $deg=1$, and $T=0.6$, correct subdivision of he variables into the global and local ones were performed in all problems.
+Then, all problems were solved successfully using the nested optimization scheme in appropriate time and number of iterations.
+However, the solving of the complicated series Gr-LQ with the use of the linear regression didn't give good results because of the presence of the parabolic-type variables, for which the percentages of similarity to the linear model were extremely low.
+In this connection, other parameters of the method $K=100$, $deg=2$ (the quadratic regression), and $T=0.95$ were selected to conduct the experiments.
+As a result, all parameters of the problem were subdivided into the local and global ones correctly, and all problems were solved successfully.
+% in appropriate time and number of iterations.
 
-Численные результаты серий экспериментов приведены в Таблице \ref{tab1}.
-В ней представлены данные о числе решенных задач $S$, среднем числе итераций глобального поиска на верхнем уровне рекурсии $G_{av}$, среднем числе испытаний на нижнем уровне $L_{av}$, среднем времени решения задачи $t_{av}$ (в секундах).
-Вычисления проводились на компьютере с CPU Intel Core™ i7 10750H, 2.6 GHz.
-Для построения и анализа регрессионной модели использовался пакет scikit-learn из Python; global search algorithm and nested optimization scheme реализованы на C++.
+In the third series of experiments, the whole series of the problems was not solved successfully in appropriate time and number of iterations because of the specific features of the construction of the GKLS functions.
+Most variables of the GKLS functions in different sections are strongly similar to the quadratic functions. 
+Therefore, one has to raise the threshold $T$ to exclude an accidental discarding a global variable.
+As a consequence, some local parameters were classified as the global ones as well.
+A large number of global variables contribute to the increase in the number of iterations and in the problem solving time naturally.
+On the other hand, if one sets the threshold for including a variable into the local functions low enough, the method wouldn't find the global variables in some functions at all.
+
+The numerical results of the series of experiments are presented in Table \ref{tab1}.
+In this one, the data on the number of problems solved $S$, the averaged number of iterations of the global search at the upper nesting level $G_{av}$, the averaged number of trials at the lower nesting level $L_{av}$, and the averaged time of solving a single problem $t_{av}$ (in seconds) are presented.
+The computations were carried out using a computer with CPU Intel Core i7 10750H, 2.6 GHz. 
+The scikit-learn package from Python 3.9 was used for the constructing and analysis of the regression model. 
+The global search algorithm and nested optimization scheme were implemented using C++.
 
 \begin{table}[ht]
-	\caption{Результаты решения серий тестовых задач}
+	\caption{Results of solving the series of test problems}
 	\label{tab1}
@@
     \hline
-		GRIS-L & 100/100  & 847 &  451 861 & 1.4 \\
-		GRIS-LQ & 100/100 & 751 &  448 326 & 1.2 \\
+		Gr-L & 100/100  & 847 &  451 861 & 1.4 \\
+		Gr-LQ & 100/100 & 751 &  448 326 & 1.2 \\
 		GKLS-L & 18/20 & 9531 &  3 899 417 & 16.4 \\

### 2022-06-29 2022/Saransk/theses.tex
добавлена первая версия тезисов Разделение параметров

% !!!!!!!!!!!!!!!!!!   ВНИМАНИЕ !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
% Заголовки разделов формируются при помощи команд \section{}, \subsection{}, \subsubsection{}
% Не используйте уровень вложенности заголовков больше трех!
% -----------------------------
% Для оформления теорем, лемм, следствий используйте окружения 
% Def     - Определение
% Teor    - Теорема
% Lem     - Лемма
% Predl   - Предложение
% Ass     - Утверждение
% Cor     - Следствие
% Example - Пример
% -----------------------------
% Доказательство теоремы начинается командой \proof и завершается командой \endproof
% -----------------------------
% Литература помещается в окружение biblio.

\documentclass[11pt, oneside, a4paper]{article}
%\usepackage[cp1251]{inputenc} % кодировка
\usepackage[utf8]{inputenc} % кодировка
\usepackage[english, russian]{babel} % Русские и английские переносы
\usepackage{graphicx}          % для включения графических изображений
\usepackage{cite}              % для корректного оформления литературы
\usepackage{enumitem}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{mathtext}

%стилевой пакет
\usepackage{schoolseminar2022}

%\renewcommand{\thefootnote}{*}\footnote{Работа выполнена при частичной финансовой поддержке .....}

\begin{document}
% \udk     - универсальный десятичный классификатор
% \msc     - Индекс предметной классификации (Mathematics Subject Classification)
% \title   - название статьи
% \authors - список авторов
\setcounter{page}{1}

%\udk{КодУДК}

\title{Использование методов машинного обучения для разделения параметров в задачах глобальной оптимизации большой размерности\footnote{Работа выполнена при частичной финансовой поддержке .....}}

\authors{Баркалов К.А., Усова М.А.}
\organizations{Нижегородский государственный университет им. Н.И. Лобачевского, Нижний Новгород, Россия}

% \section{название} - заголовок раздела первого уровня
% \subsection{название} - заголовок раздела второго уровня
% \subsubsection{название} - заголовок раздела третьего уровня

\bigskip

В настоящее время методы глобальной оптимизации используются в различных сферах науки и техники. В реальных моделях число параметров задачи может составлять десятки и сотни. Использование детерминированных методов для решения задачах такой размерности крайне ограничено из-за чрезвычайно больших вычислительных затрат даже при использовании эффективных алгоритмов (например, \cite{Paulavicius2011,Evtushenko2009,Jones2009}). 
Многие обратные задачи характеризуются тем, что зависимость от разных групп параметров носит разный характер. Например, от одной группы параметров зависимость может быть близка к линейной, тогда как по второй группе зависимость может носить сложный многоэкстремальный характер. Такие задачи могут быть решены с помощью рекурсивной оптимизации, однако заранее указать разделение на группы параметров, как правило, не представляется возможным, т.к. целевая функция в обратных задачах задается как черный ящик.

В работе рассматривается задача оптимизации вида 
\begin{eqnarray}\label{main_problem}
& \varphi(y^\ast)=\min{\left\{\varphi(y): y\in D\right\}}, \nonumber \\
& D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
\end{eqnarray}
Будем предполагать, что функция $\varphi(y)$ является многоэкстремальной и удовлетворяет условию Липшица с априори неизвестной константой Липшица. Считается, что в данной задаче зависимость от разных групп параметров носит разный характер.

Предложена схема выделения параметров задачи, которые оказывают локальное влияние на целевую функцию, что позволяет решать существенно многомерные задачи с использованием реккурсивной оптимизации \cite{Carr}. Разделение параметров происходит в 3 этапа:

Этап 1. Зафиксировать опорную точку.

Этап 2. Провести исследование на локальность по каждой переменной: вычислить в заданном количестве точек значение целевой функции в опорной точке при варьировании исследуемой переменной; на основе полученных данных построить регрессионную модель заданной степени; вычислить оценку $R^2$ и на её основе провести классификацию переменной.

Этап 3. Если локальные переменные были обнаружены, то продолжить решение задачи с использованием многошаговой схемы с разделением локальных и глобальных переменных по уровням. В противном случае запустить алгоритм глобального поиска, считая все переменные глобальными.

В основе проводимой регрессии лежит метод наименьших квадратов (МНК). Оценка качества построенной модели на шаге 2 производится с помощью коэффициента детерминации $R^2$. Подробная схема алгоритма будет представлена в полной версии публикации.

Предложенный подход показал свою работоспособность при решении нескольких серий тестовых задач, представляющих собой линейные комбинации подзадач с глобальными параметрами (функции Гришагина или GRIS \cite{Grishagin1994}, функции Сергеева или GKLS \cite{Sergeyev2015,Grishagin2018}) и подзадач с локальными параметрами, представляющих собой комбинацию близких к линейным одномерных функций (далее L) или  комбинацию  функций с разным вкладом квадратичной составляющей (далее LQ).

По результатам экспериментов при решении серий задач с существенной многоэкстремальностью (GRIS) было проведено корректное разделение на глобальные и локальные переменные, а затем задачи были успешно решены при помощи многошаговой схемы за приемлемое время и количество итераций. Задачи с функциями  GKLS для алгоритма являются более сложными, в силу схожести части параметров данных функций с параболоидами, поэтому решить все задачи серии за ограниченное количество итераций метода не удалось.

Вычисления проводились на компьютере с CPU Intel Core™ i7 10750H, 2.6 GHz. Для построения и анализа регрессионной модели использовался пакет scikit-learn из Python.
Численные результаты серий экспериментов приведены в Таблице \ref{tab1}.
В ней представлены данные о числе решенных задач $S$, среднем числе итераций глобального поиска на верхнем уровне рекурсии $G_{av}$, среднем числе испытаний на нижнем уровне $L_{av}$, среднем времени решения задачи $t_{av}$ (в секундах).

\begin{table}[ht]
	\caption{Результаты решения серий тестовых задач}
	\label{tab1}
	\begin{center}
		\begin{tabular}{ l c c c c c c } \hline
		 & $S$ &  $G_{av}$ &  $L_{av}$ & $t_{av}$ \\
    \hline
		GRIS-L & 20/20  & 705 &  378 545 & 1.3 \\
		GRIS-LQ & 20/20 & 705 &  420 397 & 1.4 \\
		GKLS-L & 18/20 & 9531 &  3 899 417 & 16.4 \\
		\hline
		\end{tabular}
	\end{center}
\end{table}

\begin{biblio}

\bibitem{fio_bib1} Шаманаев П. А., Язовцева О. С. Достаточные условия локальной покомпонентной асимптотической эквивалентности нелинейных систем обыкновенных дифференциальных уравнений и ее приложение к устойчивости по части переменных // Журнал Средневолжского математического общества. 2017. Т. 19, № 1. С. 102--115.

\bibitem{fio_bib2} Шаманаев П. А., Язовцева О. С. Достаточные условия полиустойчивости по части переменных нулевого решения нелинейных систем обыкновенных дифференциальных уравнений // Журнал Средневолжского математического общества.  2018. Т. 20, № 3. С. 304--317.

\bibitem{fio_bib3} Шаманаев П. А., Язовцева О. С. Исследование устойчивости положения равновесия системы динамики биоценоза в условиях межвидового взаимодействия // Вестник Мордовского университета. 2018. Т. 28, № 3. С. 321–332.

\bibitem{fio_bib4} Малкин И. Г. Теория устойчивости движения. М.: Наука, 1966. 533 с.

\end{biblio}

\title{Thesis title}

\authors{FIO1 \superscript{1}, FIO2 \superscript{2}}
\organizations{ Organization1 \superscript{1}, 	Organization2 \superscript{2} }

\end{document}

### 2022-06-29 2022/Saransk/theses.tex
some corrections

@@
 
-\title{Использование методов машинного обучения для разделения параметров в задачах глобальной оптимизации большой размерности\footnote{Работа выполнена при частичной финансовой поддержке .....}}
+\title{Разделение параметров в задачах глобальной оптимизации с помощью методов машинного обучения\footnote{Работа выполнена при поддержке Министерства науки и высшего образования РФ (проект \textnumero~0729-2020-0055) и научно-образовательного математического центра <<Математика технологий будущего>> (проект \textnumero~075-02-2021-1394).}}
 
 \authors{Баркалов К.А., Усова М.А.}
-\organizations{Нижегородский государственный университет им. Н.И. Лобачевского, Нижний Новгород, Россия}
+\organizations{Нижегородский государственный университет им. Н.И. Лобачевского \\ Нижний Новгород, Россия}
 
@@
 
-В настоящее время методы глобальной оптимизации используются в различных сферах науки и техники.
-В реальных моделях число параметров задачи может составлять десятки и сотни.
-Использование детерминированных методов для решения задачах такой размерности крайне ограничено из-за чрезвычайно больших вычислительных затрат даже при использовании эффективных алгоритмов (например, \cite{Paulavicius2011,Evtushenko2009,Jones2009}).
-Многие обратные задачи характеризуются тем, что зависимость от разных групп параметров носит разный характер.
+В настоящее время методы глобальной оптимизации используются в различных сферах науки и техники, например, для решения обратных задач химической кинетики.
+В обратных задачах число параметров задачи может составлять десятки и сотни.
+Использование детерминированных методов для решения задач такой размерности крайне ограничено из-за чрезвычайно больших вычислительных затрат даже при использовании эффективных алгоритмов (например, \cite{Paulavicius2011,Evtushenko2009,Jones2009}).
+При этом многие обратные задачи характеризуются тем, что зависимость от разных групп параметров носит разный характер.
 Например, от одной группы параметров зависимость может быть близка к линейной, тогда как по второй группе зависимость может носить сложный многоэкстремальный характер.
-Такие задачи могут быть решены с помощью рекурсивной оптимизации, однако заранее указать разделение на группы параметров, как правило, не представляется возможным, т.к.
+Такие задачи могут быть решены с помощью рекурсивной оптимизации.
+Однако заранее указать разделение на группы параметров, как правило, не представляется возможным, т.к.
 целевая функция в обратных задачах задается как черный ящик.
 
-В работе рассматривается задача оптимизации вида 
+В работе рассматриваются задачи оптимизации вида 
 \begin{eqnarray}\label{main_problem}
@@
 \end{eqnarray}
-Будем предполагать, что функция $\varphi(y)$ является многоэкстремальной и удовлетворяет условию Липшица с априори неизвестной константой Липшица.
+Будем предполагать, что функция $\varphi(y)$ является многоэкстремальной и удовлетворяет условию Липшица с априори неизвестной константой.
 Считается, что в данной задаче зависимость от разных групп параметров носит разный характер.
@@
 
-\title{Thesis title}
+\title{Parameters Separation in Global Optimization Problems Using Machine Learning Methods}
 
-\authors{FIO1 \superscript{1}, FIO2 \superscript{2}}
-\organizations{ Organization1 \superscript{1}, 	Organization2 \superscript{2} }
+\authors{Konstantin Barkalov, Marina Usova}
+\organizations{ Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia}
 

### 2022-06-30 2022/Saransk/theses.tex
new references

@@
 В обратных задачах число параметров задачи может составлять десятки и сотни.
-Использование детерминированных методов для решения задач такой размерности крайне ограничено из-за чрезвычайно больших вычислительных затрат даже при использовании эффективных алгоритмов (например, \cite{Paulavicius2011,Evtushenko2009,Jones2009}).
+Использование детерминированных методов для решения задач такой размерности крайне ограничено из-за чрезвычайно больших вычислительных затрат даже при использовании эффективных алгоритмов (например, \cite{Evtushenko2009,Paulavicius2016}).
 При этом многие обратные задачи характеризуются тем, что зависимость от разных групп параметров носит разный характер.
@@
 
-Предложена схема выделения параметров задачи, которые оказывают локальное влияние на целевую функцию, что позволяет решать существенно многомерные задачи с использованием реккурсивной оптимизации \cite{Carr}.
+Предложена схема выделения параметров задачи, которые оказывают локальное влияние на целевую функцию, что позволяет решать существенно многомерные задачи с использованием рекурсивной оптимизации \cite{Grishagin2007}.
 Разделение параметров происходит в 3 этапа:
@@
 
-Предложенный подход показал свою работоспособность при решении нескольких серий тестовых задач, представляющих собой линейные комбинации подзадач с глобальными параметрами (функции Гришагина или GRIS \cite{Grishagin1994}, функции Сергеева или GKLS \cite{Sergeyev2015,Grishagin2018}) и подзадач с локальными параметрами, представляющих собой комбинацию близких к линейным одномерных функций (далее L) или  комбинацию  функций с разным вкладом квадратичной составляющей (далее LQ).
+Предложенный подход показал свою работоспособность при решении нескольких серий тестовых задач, представляющих собой линейные комбинации подзадач с глобальными параметрами (функции Гришагина или GRIS \cite{Grishagin1994}, функции Сергеева или GKLS \cite{Sergeyev2008}) и подзадач с локальными параметрами, представляющих собой комбинацию близких к линейным одномерных функций (далее L) или  комбинацию  функций с разным вкладом квадратичной составляющей (далее LQ).
 
@@
 \begin{biblio}
+
+\bibitem{Grishagin2007} Городецкий~С.Ю., Гришагин~В.А. Нелинейное программирование и многоэкстремальная оптимизация. Н.~Новгород: Изд-во ННГУ, 2007. 489~с.
+
+\bibitem{Grishagin2001}
+{\it Sergeyev Y., Grishagin V.} Parallel Asynchronous Global Search and the Nested Optimization Scheme // J. Comput. Anal. Appl. 2001. V.~3~(2). P.~123--145.
+
+\bibitem{Sergeyev2008} Sergeyev Y.D., Kvasov D.E. Diagonal'nyye metody global'noy optimizatsii. Fizmatlit, 2008. P. 352.
+
+\bibitem{Evtushenko2009}
+{\it Евтушенко Ю.Г., Малкова В.У., Станевичюс А.-И.А.} Параллельный поиск глобального экстремума функций многих переменных // Журн. вычисл. матем. и матем. физ. 2009. Т.~49. №~2. C.~255--269.
+
+\bibitem{Paulavicius2016}
+{\it Paulavi{\v c}ius R., {\v Z}ilinskas J.} Advantages of Simplicial Partitioning for Lipschitz Optimization Problems with Linear Constraints // Optim. Lett. 2016. V.~10~(2). P.~237--246.
+
+\bibitem{Strongin13}
+{\it Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Изд-во МГУ, 2013.
 

### 2022-06-30 2022/OPTIMA/paper.tex

--- 2022-01-14 2022/OM&S/paper.tex
@@
-% interacttfssample.tex
-% v1.05 - August 2017
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.21 of 2022/01/12
+%
+\documentclass[runningheads]{llncs}
 
-\documentclass[]{interact}
+\usepackage{amsmath}
+%
+\usepackage[T1]{fontenc}
+% T1 fonts will be used to generate the final print and online PDFs,
+% so please use T1 fonts in your manuscript whenever possible.
+% Other font encondings may result in incorrect characters.
+%
+\usepackage{graphicx}
 
-\usepackage{color}
-\usepackage{hyperref}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following two lines
+% to display URLs in blue roman font according to Springer's eBook style:
+%\usepackage{color}
+%\renewcommand\UrlFont{\color{blue}\rmfamily}
+%
+\begin{document}
+%
+\title{An accelerated algorithm for finding efficient solutions in multiobjective problems with black-box multiextremal criteria\thanks{This work was supported by the Russian Science Foundation, project No.
+21-11-00204.}}
+%
+\titlerunning{Algorithm for finding efficient solutions in multiobjective problems}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
+%
+\author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
+Vladimir Grishagin\orcidID{0000-0002-2884-3670} \and
+Evgeny Kozinov\orcidID{0000-0001-6776-0096}}
+%
+\authorrunning{K. Barkalov, V. Grishagin and E. Kozinov}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, \email{vagris@unn.ru}}
 
-\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
-\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
-%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
-
-%\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
-%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled
-%\setlength\bibindent{2em}% To increase hanging indent in bibliography when line spacing is doubled
-
-\usepackage[numbers,sort&compress]{natbib}% Citation support using natbib.sty
-\bibpunct[, ]{[}{]}{,}{n}{,}{,}% Citation support using natbib.sty
-\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty
-
-\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{proposition}[theorem]{Proposition}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-
-\theoremstyle{remark}
-\newtheorem{remark}{Remark}
-\newtheorem{notation}{Notation}
-
-\begin{document}
-
-%\articletype{ARTICLE TEMPLATE}% Specify the article type or omit as appropriate
-
-\title{An accelerated algorithm for finding efficient solutions in multiobjective problems with black-box multiextremal criteria}
-
-\author{
-\name{Konstantin Barkalov, Vladimir Grishagin and Evgeny Kozinov\thanks{CONTACT Konstantin Barkalov. Email: konstantin.barkalov@itmm.unn.ru}}
-\affil{Institute of Information Technology, Mathematics and Mechanics, Lobachevsky State University, 23 Gagarin ave., 603022 Nizhny Novgorod, Russia}
-}
-
-\maketitle
-
+%
+\maketitle              % typeset the header of the contribution
+%
 \begin{abstract}
@@
 Comparison with 5 nature-inspired particle swarm and genetic algorithms on the base of the hypervolume index reflecting quality of Pareto set evaluation demonstrates qualitative results of the proposed method.
+
+\keywords{Multiobjective optimization \and Global optimization \and Dimensionality reduction \and Search information \and Nature-inspired metaheuristics \and Numerical comparison.}
 \end{abstract}
-
-\begin{keywords}
-Multiobjective optimization; global optimization; dimensionality reduction; search information; nature-inspired metaheuristics; numerical comparison
-\end{keywords}
-
+%
+%
+%
 \section{Introduction}
@@
 Here
-\begin{gather}\label{eq:18}
+\begin{equation}\label{eq:18}
+\begin{matrix}
 B(q) = (t_s - t_{s-1})^{1/N} + \frac{(v_q - v_{q-1})^2} {m^2 (t_s - t_{s-1})^{1/N}} -2\frac{v_{q} + v_{q-1}}{m},\\
 \varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
-\end{gather}
+\end{matrix}
+\end{equation}
 is minimal computed value of the objective function $\varphi(t)$ at trial points,
@@
 \begin{table}
-\tbl{Results of the experiment with bi-criteria MCO problems of dimension 2.}
-{\begin{tabular}{ccccccc} \toprule
- Pareto & RALR-DA & OMOPSO & SMPSO & NSGAII & IBEA & SPEA2 \\ \midrule
-  6.62 & 6.44 & 5.93 & 5.88 & 5.71 & 5.65 & 5.61 \\ \bottomrule
-\end{tabular}}
+\caption{Results of the experiment with bi-criteria MCO problems of dimension 2.}
+\centering
+\begin{tabular}{ccccccc} \hline
+ Pareto & RALR-DA & OMOPSO & SMPSO & NSGAII & IBEA & SPEA2 \\ \hline
+  6.62 & 6.44 & 5.93 & 5.88 & 5.71 & 5.65 & 5.61 \\ \hline
+\end{tabular}
 \label{tab1}
@@
 \begin{table}
-\tbl{Results of testing bi-criteria 4-dimensional MCO problems.}
-{\begin{tabular}{ccccccc} \toprule
- Pareto & RALR-DA & SPEA2 & OMOPSO & IBEA & NSGAII & SMPSO   \\ \midrule
-  5.04 & 4.71 & 4.53 & 4.45 & 4.39 & 4.39 & 4.37 \\ \bottomrule
-\end{tabular}}
+\caption{Results of testing bi-criteria 4-dimensional MCO problems.}
+\centering
+\begin{tabular}{ccccccc} \hline
+ Pareto & RALR-DA & SPEA2 & OMOPSO & IBEA & NSGAII & SMPSO   \\ \hline
+  5.04 & 4.71 & 4.53 & 4.45 & 4.39 & 4.39 & 4.37 \\ \hline
+\end{tabular}
 \label{tab2}
@@
 \begin{table}
-\tbl{Results of testing 2-dimensional MCO problems with 4 criteria.}
-{\begin{tabular}{ccccccc} \toprule
- Pareto & RALR-DA & OMOPSO & SPEA2 & SMPSO & IBEA & NSGAII   \\ \midrule
-  23.89 & 23.52 & 22.87 & 22.80 & 22.53 & 20.68 & 19.82 \\ \bottomrule
-\end{tabular}}
+\caption{Results of testing 2-dimensional MCO problems with 4 criteria.}
+\centering
+\begin{tabular}{ccccccc} \hline
+ Pareto & RALR-DA & OMOPSO & SPEA2 & SMPSO & IBEA & NSGAII   \\ \hline
+  23.89 & 23.52 & 22.87 & 22.80 & 22.53 & 20.68 & 19.82 \\ \hline
+\end{tabular}
 \label{tab3}
@@
 
-\section*{Disclosure statement}
-
-No potential conflict of interest was reported by the authors.
-
-\section*{Funding}
-
-This work was supported by the Russian Science Foundation, project No. 21-11-00204.
-
-\bibliographystyle{tfs}
+\bibliographystyle{splncs04}
 \bibliography{bibliography}
@@
 \end{document}
-+

### 2022-06-30 2022/OPTIMA/paper.tex
reduce number of references, 16 pages now

@@
 
-Note that special classes of multiextremal optimization problems arise when the criteria can be described analytically.
+%Note that special classes of multiextremal optimization problems arise when the criteria can be described analytically.
 In this case one can apply efficient methods based on d.c.
@@
 As the directions of further development of the current research it would be interesting to investigate the possibilities of applying other schemes of dimensionality reduction, for example,
-the nested optimization scheme \cite{Grishagin2016,Grishagin2018}, the diagonal \cite{Sergeyev2015,Sergeyev2017} or simplicial \cite{Zilinskas2011,PaulaviciusZilinskas2014} partition methods
+the nested optimization scheme \cite{Grishagin2016,Grishagin2018}, the diagonal \cite{Sergeyev2017} or simplicial \cite{PaulaviciusZilinskas2014} partition methods
 in the framework of the general approach used in the paper. Moreover, the prospective way is to develop parallel versions of the proposed general scheme oriented at powerful supercomputer systems.

### 2022-06-30 2022/Saransk/theses.tex

@@
 Этап 2.
-Провести исследование на локальность по каждой переменной: вычислить в заданном количестве точек значение целевой функции в опорной точке при варьировании исследуемой переменной; на основе полученных данных построить регрессионную модель заданной степени; вычислить оценку $R^2$ и на её основе провести классификацию переменной.
+Провести исследование на локальность по каждой переменной: вычислить в заданном количестве точек значение целевой функции в опорной точке при варьировании исследуемой переменной; на основе полученных данных построить регрессионную модель; вычислить оценку $R^2$ и на её основе провести классификацию переменной.
 
@@
 Если локальные переменные были обнаружены, то продолжить решение задачи с использованием многошаговой схемы с разделением локальных и глобальных переменных по уровням.
-В противном случае запустить алгоритм глобального поиска, считая все переменные глобальными.
+В противном случае запустить алгоритм глобального поиска \cite{Grishagin2007}, считая все переменные глобальными.
 
@@
 
-Предложенный подход показал свою работоспособность при решении нескольких серий тестовых задач, представляющих собой линейные комбинации подзадач с глобальными параметрами (функции Гришагина или GRIS \cite{Grishagin1994}, функции Сергеева или GKLS \cite{Sergeyev2008}) и подзадач с локальными параметрами, представляющих собой комбинацию близких к линейным одномерных функций (далее L) или  комбинацию  функций с разным вкладом квадратичной составляющей (далее LQ).
+Предложенный подход показал свою работоспособность при решении нескольких серий тестовых задач, представляющих собой линейные комбинации подзадач с глобальными параметрами (функции Гришагина или GRIS, функции Сергеева или GKLS \cite{Grishagin2001}) и подзадач с локальными параметрами, представляющих собой комбинацию близких к линейным одномерных функций (далее L) или  комбинацию  функций с разным вкладом квадратичной составляющей (далее LQ).
 
 По результатам экспериментов при решении серий задач с существенной многоэкстремальностью (GRIS) было проведено корректное разделение на глобальные и локальные переменные, а затем задачи были успешно решены при помощи многошаговой схемы за приемлемое время и количество итераций.
-Задачи с функциями  GKLS для алгоритма являются более сложными, в силу схожести части параметров данных функций с параболоидами, поэтому решить все задачи серии за ограниченное количество итераций метода не удалось.
+Задачи с функциями  GKLS для алгоритма являются более сложными, в силу схожести части параметров функций с параболоидами, поэтому решить все задачи серии за ограниченное количество итераций метода не удалось.
 
@@
 
-\bibitem{Grishagin2007} Городецкий~С.Ю., Гришагин~В.А. Нелинейное программирование и многоэкстремальная оптимизация. Н.~Новгород: Изд-во ННГУ, 2007. 489~с.
+\bibitem{Evtushenko2009} Евтушенко Ю.
+Г., Малкова В.
+У., Станевичюс А.-И.
+А.
+Параллельный поиск глобального экстремума функций многих переменных // Журн.
+вычисл.
+матем.
+и матем.
+физ.
+2009.
+Т.
+49.
+№ 2.
+C.
+255--269.
 
-\bibitem{Grishagin2001}
-{\it Sergeyev Y., Grishagin V.} Parallel Asynchronous Global Search and the Nested Optimization Scheme // J. Comput. Anal. Appl. 2001. V.~3~(2). P.~123--145.
+\bibitem{Paulavicius2016} Paulavi{\v c}ius R., {\v Z}ilinskas J.
+Advantages of Simplicial Partitioning for Lipschitz Optimization Problems with Linear Constraints // Optim.
+Lett.
+2016.
+V.
+10.
+No.
+2.
+P.
+237--246.
 
-\bibitem{Sergeyev2008} Sergeyev Y.D., Kvasov D.E. Diagonal'nyye metody global'noy optimizatsii. Fizmatlit, 2008. P. 352.
+\bibitem{Grishagin2007} Городецкий С. Ю., Гришагин В. А. Нелинейное программирование и многоэкстремальная оптимизация. Н. Новгород: Изд-во ННГУ, 2007. 489 с.
 
-\bibitem{Evtushenko2009}
-{\it Евтушенко Ю.Г., Малкова В.У., Станевичюс А.-И.А.} Параллельный поиск глобального экстремума функций многих переменных // Журн. вычисл. матем. и матем. физ. 2009. Т.~49. №~2. C.~255--269.
+\bibitem{Grishagin2001} Sergeyev Y., Grishagin V. Parallel Asynchronous Global Search and the Nested Optimization Scheme // J. Comput. Anal. Appl. 2001. V. 3, No. 2. P. 123--145.
 
-\bibitem{Paulavicius2016}
-{\it Paulavi{\v c}ius R., {\v Z}ilinskas J.} Advantages of Simplicial Partitioning for Lipschitz Optimization Problems with Linear Constraints // Optim. Lett. 2016. V.~10~(2). P.~237--246.
+%\bibitem{Sergeyev2008} Сергеев Я. Д., Квасов Д. Е. Диагональные методы глобальной оптимизации. М.: Физматлит, 2008. 352 с.
 
-\bibitem{Strongin13}
-{\it Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Изд-во МГУ, 2013.
-
-\bibitem{fio_bib1} Шаманаев П.
-А., Язовцева О.
-С.
-Достаточные условия локальной покомпонентной асимптотической эквивалентности нелинейных систем обыкновенных дифференциальных уравнений и ее приложение к устойчивости по части переменных // Журнал Средневолжского математического общества.
-2017.
-Т.
-19, № 1.
-С.
-102--115.
-
-\bibitem{fio_bib2} Шаманаев П.
-А., Язовцева О.
-С.
-Достаточные условия полиустойчивости по части переменных нулевого решения нелинейных систем обыкновенных дифференциальных уравнений // Журнал Средневолжского математического общества.
-2018.
-Т.
-20, № 3.
-С.
-304--317.
-
-\bibitem{fio_bib3} Шаманаев П.
-А., Язовцева О.
-С.
-Исследование устойчивости положения равновесия системы динамики биоценоза в условиях межвидового взаимодействия // Вестник Мордовского университета.
-2018.
-Т.
-28, № 3.
-С.
-321–332.
-
-\bibitem{fio_bib4} Малкин И. Г. Теория устойчивости движения. М.: Наука, 1966. 533 с.
+%\bibitem{Strongin13} Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А. Параллельные вычисления в задачах глобальной оптимизации. М.: Изд-во МГУ, 2013. 280 с.
 

### 2022-07-01 2022/OPTIMA/paper.tex
Обедилин рисунки и таблицы

@@
 Here Fig.~\ref{fig2a} corresponds to the Pareto front computed on the uniform grid with the step 0.05 in the search domain whereas Fig.~\ref{fig2b} shows the Pareto front computed by RALR-DA.
+The averaged HV indices obtained while solving the two-dimensional bi-criteria problems are presented in Tab.~\ref{tab4_6} (series 1) and Fig.~\ref{fig4_6} where ``Pareto'' corresponds to the estimation obtained by the grid method.
 
@@
 
-The averaged HV indices obtained while solving the two-dimensional bi-criteria problems are presented in Tab.~\ref{tab1} and Fig.~\ref{fig3} where ``Pareto'' corresponds to the estimation obtained by the grid method.
+The second experiment was aimed at solving bi-criteria 4-dimensional MCO problems.
+10 test problems have been investigated and for each of them, 100 convolutions (\ref{conv_problem}) with uniformly distributed weight coefficients $\lambda$ were minimized by RALR-DA.
+Tab.~\ref{tab4_6} (series 2) contains the results achieved by the methods participated in the experiment and Fig.~\ref{fig4_6} presents them in graphical form.
+
+The third collection of test problems consisted of 10 two-dimensional MCO benchmarks with 4 criteria.
+In this experiment RALR-DA in each MCO problem minimized convolutions (\ref{conv_problem}) corresponding to 100 different values of weight coefficients $\lambda$ uniformly distributed in the domain (\ref{lambda}).
+Like  the  previous experiments, the results are presented in Tab.~\ref{tab4_6} (series 3) and Fig.~\ref{fig4_6}.
 
 \begin{table}
-\caption{Results of the experiment with bi-criteria MCO problems of dimension 2.}
+\caption{Results of the experiment with multidimensional MCO problems.}
 \centering
-\begin{tabular}{ccccccc} \hline
- Pareto & RALR-DA & OMOPSO & SMPSO & NSGAII & IBEA & SPEA2 \\ \hline
-  6.62 & 6.44 & 5.93 & 5.88 & 5.71 & 5.65 & 5.61 \\ \hline
+\begin{tabular}{cccccccccc} \hline
+Series& N& k& Pareto & RALR-DA & OMOPSO & SPEA2 & SMPSO & IBEA & NSGAII   \\ \hline
+1& 2&	2 & 6.62 & 6.44 & 5.93 & 5.88 & 5.71 & 5.65 & 5.61 \\ 
+2& 4&	2 &  5.04 & 4.71 & 4.53 & 4.45 & 4.39 & 4.39 & 4.37 \\ 
+3& 2&	4 &  23.89 & 23.52 & 22.87 & 22.80 & 22.53 & 20.68 & 19.82 \\ \hline
 \end{tabular}
-\label{tab1}
+\label{tab4_6}
 \end{table}
@@
 \centering
-\includegraphics[width=0.7\textwidth]{fig3.pdf} 
-\caption{HV indices in bi-criteria 2-dimensional case.}\label{fig3} 
-\end{figure}
-
-The second experiment was aimed at solving bi-criteria 4-dimensional MCO problems.
-10 test problems have been investigated and for each of them, 100 convolutions (\ref{conv_problem}) with uniformly distributed weight coefficients $\lambda$ were minimized by RALR-DA.
-Tab.~\ref{tab2} contains the results achieved by the methods participated in the experiment and Fig.~\ref{fig4} presents them in graphical form.
-
-\begin{table}
-\caption{Results of testing bi-criteria 4-dimensional MCO problems.}
-\centering
-\begin{tabular}{ccccccc} \hline
- Pareto & RALR-DA & SPEA2 & OMOPSO & IBEA & NSGAII & SMPSO   \\ \hline
-  5.04 & 4.71 & 4.53 & 4.45 & 4.39 & 4.39 & 4.37 \\ \hline
-\end{tabular}
-\label{tab2}
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.7\textwidth]{fig4.pdf} 
-\caption{HV indices in bi-criteria 4-dimensional case.}\label{fig4} 
-\end{figure}
-
-The third collection of test problems consisted of 10 two-dimensional MCO benchmarks with 4 criteria.
-In this experiment RALR-DA in each MCO problem minimized convolutions (\ref{conv_problem}) corresponding to 100 different values of weight coefficients $\lambda$ uniformly distributed in the domain (\ref{lambda}).
-Like  the  previous experiments, the results are presented in Tab.~\ref{tab3} and Fig.~\ref{fig5}.
-
-\begin{table}
-\caption{Results of testing 2-dimensional MCO problems with 4 criteria.}
-\centering
-\begin{tabular}{ccccccc} \hline
- Pareto & RALR-DA & OMOPSO & SPEA2 & SMPSO & IBEA & NSGAII   \\ \hline
-  23.89 & 23.52 & 22.87 & 22.80 & 22.53 & 20.68 & 19.82 \\ \hline
-\end{tabular}
-\label{tab3}
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.7\textwidth]{fig5.pdf} 
-\caption{HV indices in 4-criteria 2-dimensional case.}\label{fig5} 
+\includegraphics[width=0.7\textwidth]{fig4_6.pdf} 
+\caption{Relative backlog of the HV index ($HV_{Pareto} / HV_{alg}$).}\label{fig4_6} 
 \end{figure}

### 2022-07-01 2022/OPTIMA/paper.tex
fix

@@
 \usepackage{amsmath}
+\usepackage[dvipsnames]{xcolor}
 %
@@
 
-In the first experiment 100 two-dimensional bi-criteria problems have been taken.
+In the first experiment 100 two-dimensional bi-criteria problems have been taken \textcolor{red}{($N=2$, $k=2$)}.
 The contour plots of a convolution $\Phi_\lambda(x)$ from (\ref{conv_problem})  demonstrating its multiextremal and non-smooth relief are presented in Fig.~\ref{fig1}.
@@
 Here Fig.~\ref{fig2a} corresponds to the Pareto front computed on the uniform grid with the step 0.05 in the search domain whereas Fig.~\ref{fig2b} shows the Pareto front computed by RALR-DA.
-The averaged HV indices obtained while solving the two-dimensional bi-criteria problems are presented in Tab.~\ref{tab4_6} (series 1) and Fig.~\ref{fig4_6} where ``Pareto'' corresponds to the estimation obtained by the grid method.
+The averaged HV indices obtained while solving the two-dimensional bi-criteria problems are presented in \textcolor{red}{ Tab.~\ref{tab4_6} (series 1) where column $N$ is the dimension of the MCO problems, column $k$ corresponds to criteria counts and column ``Pareto'' corresponds to the estimation obtained by the grid method.}
 
@@
 
-The second experiment was aimed at solving bi-criteria 4-dimensional MCO problems.
-10 test problems have been investigated and for each of them, 100 convolutions (\ref{conv_problem}) with uniformly distributed weight coefficients $\lambda$ were minimized by RALR-DA.
-Tab.~\ref{tab4_6} (series 2) contains the results achieved by the methods participated in the experiment and Fig.~\ref{fig4_6} presents them in graphical form.
-
-The third collection of test problems consisted of 10 two-dimensional MCO benchmarks with 4 criteria.
-In this experiment RALR-DA in each MCO problem minimized convolutions (\ref{conv_problem}) corresponding to 100 different values of weight coefficients $\lambda$ uniformly distributed in the domain (\ref{lambda}).
-Like  the  previous experiments, the results are presented in Tab.~\ref{tab4_6} (series 3) and Fig.~\ref{fig4_6}.
-
 \begin{table}
-\caption{Results of the experiment with multidimensional MCO problems.}
+\caption{\textcolor{red}{ Results of the experiment with multidimensional MCO problems.}}
 \centering
 \begin{tabular}{cccccccccc} \hline
-Series& N& k& Pareto & RALR-DA & OMOPSO & SPEA2 & SMPSO & IBEA & NSGAII   \\ \hline
+Series& $N$& $k$& Pareto & RALR-DA & OMOPSO & SPEA2 & SMPSO & IBEA & NSGAII   \\ \hline
 1& 2&	2 & 6.62 & 6.44 & 5.93 & 5.88 & 5.71 & 5.65 & 5.61 \\ 
@@
 
+The second experiment was aimed at solving bi-criteria 4-dimensional MCO problems \textcolor{red}{($N=4$, $k=2$)}.
+10 test problems have been investigated and for each of them, 100 convolutions (\ref{conv_problem}) with uniformly distributed weight coefficients $\lambda$ were minimized by RALR-DA.
+Tab.~\ref{tab4_6} (series 2) contains the results achieved by the methods participated in the experiment.
+
+The third collection of test problems consisted of 10 two-dimensional MCO benchmarks with 4 criteria \textcolor{red}{($N=2$, $k=4$)}.
+In this experiment RALR-DA in each MCO problem minimized convolutions (\ref{conv_problem}) corresponding to 100 different values of weight coefficients $\lambda$ uniformly distributed in the domain (\ref{lambda}).
+Like  the  previous experiments, the results are presented in Tab.~\ref{tab4_6} (series 3).
+
 \begin{figure}
@@
 \includegraphics[width=0.7\textwidth]{fig4_6.pdf} 
-\caption{Relative backlog of the HV index ($HV_{Pareto} / HV_{alg}$).}\label{fig4_6} 
+\caption{\textcolor{red}{ Relative backlog of the HV index ($HV_{Pareto} / HV_{alg}$). }}\label{fig4_6} 
 \end{figure}
+
+\textcolor{red}{Fig.~\ref{fig4_6} shows the ratio of the HV indices of the Pareto set to the HV index of a specific algorithm.}
 

### 2022-07-01 2022/OPTIMA/paper.tex
fix table

@@
 \begin{tabular}{cccccccccc} \hline
-Series& $N$& $k$& Pareto & RALR-DA & OMOPSO & SPEA2 & SMPSO & IBEA & NSGAII   \\ \hline
+Series& $N$& $k$& Pareto & RALR-DA & OMOPSO &SMPSO & NSGAII & IBEA & SPEA2  \\ \hline
 1& 2&	2 & 6.62 & 6.44 & 5.93 & 5.88 & 5.71 & 5.65 & 5.61 \\ 
-2& 4&	2 &  5.04 & 4.71 & 4.53 & 4.45 & 4.39 & 4.39 & 4.37 \\ 
-3& 2&	4 &  23.89 & 23.52 & 22.87 & 22.80 & 22.53 & 20.68 & 19.82 \\ \hline
+2& 4&	2 & 5.04 & 4.71 & 4.45 & 4.37 & 4.39 & 4.39 & 4.53 \\ 
+3& 2&	4 &  23.89 & 23.52 & 22.87 & 22.53 & 19.82 & 20.68 & 22.80 \\ \hline
 \end{tabular}

### 2022-07-02 2022/OPTIMA/paper.tex
fix

@@
 Here Fig.~\ref{fig2a} corresponds to the Pareto front computed on the uniform grid with the step 0.05 in the search domain whereas Fig.~\ref{fig2b} shows the Pareto front computed by RALR-DA.
-The averaged HV indices obtained while solving the two-dimensional bi-criteria problems are presented in \textcolor{red}{ Tab.~\ref{tab4_6} (series 1) where column $N$ is the dimension of the MCO problems, column $k$ corresponds to criteria counts and column ``Pareto'' corresponds to the estimation obtained by the grid method.}
+The averaged HV indices obtained while solving the two-dimensional bi-criteria problems are presented in \textcolor{red}{ Tab.~\ref{tab4_6} (series 1) where column $N$ is the dimension of the MCO problems, column $k$ corresponds to criteria numbers and column ``Pareto'' corresponds to the estimation obtained by the grid method.}
 
@@
 \includegraphics[width=0.7\textwidth]{fig4_6.pdf} 
-\caption{\textcolor{red}{ Relative backlog of the HV index ($HV_{Pareto} / HV_{alg}$). }}\label{fig4_6} 
+\caption{\textcolor{red}{ Relative backlog of the HV index ($HV_{alg} / HV_{Pareto} $). }}\label{fig4_6} 
 \end{figure}
 
-\textcolor{red}{Fig.~\ref{fig4_6} shows the ratio of the HV indices of the Pareto set to the HV index of a specific algorithm.}
+\textcolor{red}{Fig.~\ref{fig4_6} shows the ratio of the HV index of a specific algorithm to the HV index of Pareto set.}
 

### 2022-07-07 2022/ICNAAM/ICNAAM.tex

@@
 % Title portion
-\title{Using Machine Learning for Subdivision of Parameters in Global Optimization Problems of Large Dimensionality}
+\title{Using Machine Learning for Separation of Parameters in Global Optimization Problems of Large Dimensionality}
 
@@
 
-At present, the global optimization methods are applied to solving various problems arising in various fields of science and technology. 
+At present, the global optimization methods are applied to solving various problems arising in various fields of science and technology \cite{Pinter2006}. 
 For example, the evaluation of the model parameters from the experimental data is a traditional field of application of the global optimization methods.
-Among these ones, there are, for example, the problems of chemical kinetics \cite{Gubaydullin2021}.
+Among these ones, there are the problems of chemical kinetics \cite{Gubaydullin2021}.
 In the problems of this kind, it is necessary to perform a search for the values of unknown model parameters, at which the modeling results are the closest to 
@@
 So far, the development of the algorithm for the minimization of the essentially multiextremal functions taking into account different character of the dependence of the objective function on different groups of parameters is relevant.
-In the present work, a special method for such a subdivision based on the analysis of the regression model of the function with respect to each variable separately has been proposed.
+In the present work, a special method for such a separation based on the analysis of the regression model of the function with respect to each variable separately has been proposed.
 To confirm the efficiency of the scheme proposed, a representative set of test problems has been solved.
@@
 At the same time, the subdivision of the parameters is unknown. 
-In the present work, an approach to the subdivision of the problem parameters into the two specified groups is proposed.
+In the present work, an approach to the separation of the problem parameters into the two specified groups is proposed.
 The details of the method are described in the next section.
@@
 This algorithm was demonstrated to be a powerful tool for solving complex multiextremal problems \cite{Kvasov2013,Kalyulin2017,Cavoretto2021}.
-To solve the local problem (\ref{local_problem}), we used Hooke-Jeevse pattern search method (see, e.g., \cite{Kelley}). 
+To solve the local problem (\ref{local_problem}), we used Hooke-Jeeves (pattern search) method (see, e.g., \cite{Kelley}). 
 Unlike the BFGS-type methods (see, e.g., \cite{Nocedal}), this algorithm doesn't require computing the gradient and works well under conditions of the 
@@
 
-\section{SUBDIVISION OF THE PROBLEMS PARAMETERS}
+\section{SEPARATION OF THE PROBLEMS PARAMETERS}
 
@@
 \item the point is fixed according to the physical meaning of the applied problem being solved;
-\item use the solution found by the local method in $K$ iterations (in the present study, Hooke-Jeevse pattern search method was used) as the reference point.
-\item use the solution found by the global method in $K$ iterations (in the present study, the global search algorithm was used) as the reference point.
+\item use the solution found by the local method (e.g. Hooke-Jeeves method) in $K$ iterations as the reference point;
+\item use the solution found by the global method (e.g. global search algorithm) in $K$ iterations as the reference point.
 \end{enumerate}
 
-Step 2. For all components of $y_i, \; 1\leq i \leq N$, investigate the locality:
+Step 2. For all components $y_i, \; 1\leq i \leq N$, investigate the locality:
 \begin{enumerate}
-\item compute $P+1$ values of the objective function $\varphi(y)$ at the points $z_i^j = (\overline{y}_1,...,\overline{y}_{i-1},y_i^j,\overline{y}_{i+1},...,\overline{y}_N)$ where $y_i^j =  a_i + jh, \; h=(b_i-a_i)/P, \; 0\leq j \leq P$;
+\item compute $P+1$ values of the objective function $\varphi(y)$ at the points $z_i^j = (\overline{y}_1,...,\overline{y}_{i-1},y_i^j,\overline{y}_{i+1},...,\overline{y}_N)$, where \mbox{$y_i^j = a_i + jh$},  $h=(b_i-a_i)/P, \; 0\leq j \leq P$;
 \item form a set $Q$ comprising of $P+1$ pairs of the kind $\left\{y_i^j, \varphi(z_i^j)\right\} $ 
@@
 \begin{itemize}
-\item if $R^2$ is greater than a predefined threshold $T$, then add the variable to the set of the local variables,
+\item if $R^2$ is greater than a predefined threshold $T$, then add the variable to the set of the local variables;
 \item else add the variable to the set of the global variables.
@@
 The value of $R^2_a$, as a rule, falls between 0 and 1.
-A result close to 1 shows the accurasy of the model to be high.
+A result close to 1 shows the accuracy of the model to be high.
 
@@
 
-As a result of solving the Gr-L series with the preset precision of $0.01$ and the parameters of the method $K=100$, $deg=1$, and $T=0.6$, correct subdivision of he variables into the global and local ones were performed in all problems.
+As a result of solving the Gr-L series with the preset precision of $0.01$ and the parameters of the method $K=100$, $deg=1$, and $T=0.6$, correct separation of the variables into the global and local ones were performed in all problems.
 Then, all problems were solved successfully using the nested optimization scheme in appropriate time and number of iterations.

### 2022-07-07 2022/OPTIMA/paper.tex

@@
 %
-\title{An accelerated algorithm for finding efficient solutions in multiobjective problems with black-box multiextremal criteria\thanks{This work was supported by the Russian Science Foundation, project No.
-21-11-00204.}}
+\title{An accelerated algorithm for finding efficient solutions in multiobjective problems with black-box multiextremal criteria\thanks{The work was supported by the Ministry of Science and Higher Education of the Russian Federation (project no.0729-2020-0055) and by the Research and Education Mathematical Center (project no.
+075-02-2020-1483/1).}}
 %
@@
 
-In the first experiment 100 two-dimensional bi-criteria problems have been taken \textcolor{red}{($N=2$, $k=2$)}.
+In the first experiment 100 two-dimensional bi-criteria problems have been taken ($N=2$, $k=2$).
 The contour plots of a convolution $\Phi_\lambda(x)$ from (\ref{conv_problem})  demonstrating its multiextremal and non-smooth relief are presented in Fig.~\ref{fig1}.
@@
 Here Fig.~\ref{fig2a} corresponds to the Pareto front computed on the uniform grid with the step 0.05 in the search domain whereas Fig.~\ref{fig2b} shows the Pareto front computed by RALR-DA.
-The averaged HV indices obtained while solving the two-dimensional bi-criteria problems are presented in \textcolor{red}{ Tab.~\ref{tab4_6} (series 1) where column $N$ is the dimension of the MCO problems, column $k$ corresponds to criteria numbers and column ``Pareto'' corresponds to the estimation obtained by the grid method.}
+The averaged HV indices obtained while solving the two-dimensional bi-criteria problems are presented in Tab.~\ref{tab4_6} (series 1) where column $N$ is the dimension of the MCO problems, column $k$ corresponds to criteria numbers and column ``Pareto'' corresponds to the estimation obtained by the grid method.
 
@@
 \begin{table}
-\caption{\textcolor{red}{ Results of the experiment with multidimensional MCO problems.}}
+\caption{Results of the experiment with multidimensional MCO problems.}
 \centering
@@
 
-The second experiment was aimed at solving bi-criteria 4-dimensional MCO problems \textcolor{red}{($N=4$, $k=2$)}.
+The second experiment was aimed at solving bi-criteria 4-dimensional MCO problems ($N=4$, $k=2$).
 10 test problems have been investigated and for each of them, 100 convolutions (\ref{conv_problem}) with uniformly distributed weight coefficients $\lambda$ were minimized by RALR-DA.
@@
 
-The third collection of test problems consisted of 10 two-dimensional MCO benchmarks with 4 criteria \textcolor{red}{($N=2$, $k=4$)}.
+The third collection of test problems consisted of 10 two-dimensional MCO benchmarks with 4 criteria ($N=2$, $k=4$).
 In this experiment RALR-DA in each MCO problem minimized convolutions (\ref{conv_problem}) corresponding to 100 different values of weight coefficients $\lambda$ uniformly distributed in the domain (\ref{lambda}).
@@
 \includegraphics[width=0.7\textwidth]{fig4_6.pdf} 
-\caption{\textcolor{red}{ Relative backlog of the HV index ($HV_{alg} / HV_{Pareto} $). }}\label{fig4_6} 
+\caption{Relative backlog of the HV index ($HV_{alg} / HV_{Pareto} $).}\label{fig4_6} 
 \end{figure}
 
-\textcolor{red}{Fig.~\ref{fig4_6} shows the ratio of the HV index of a specific algorithm to the HV index of Pareto set.}
+Fig.~\ref{fig4_6} shows the ratio of the HV index of a specific algorithm to the HV index of Pareto set.
 

### 2022-07-16 2022/ICNAAM/ICNAAM.tex
before proofreading

@@
 
-At present, the global optimization methods are applied to solving various problems arising in various fields of science and technology \cite{Pinter2006}. 
+At present, the global optimization methods are applied to solving various problems arising in various fields of science and technology (see \cite{Pinter2006}). 
 For example, the evaluation of the model parameters from the experimental data is a traditional field of application of the global optimization methods.
-Among these ones, there are the problems of chemical kinetics \cite{Gubaydullin2021}.
+Among these ones, there are the problems of chemical kinetics (see \cite{Gubaydullin2021}).
 In the problems of this kind, it is necessary to perform a search for the values of unknown model parameters, at which the modeling results are the closest to 
@@
 The use of the deterministic global optimization methods for solving the problems of such a dimensionality is very limited because of very large computation costs of covering the search domain by the trial points.
-This holds even in the case of using the efficient algorithms (for example, \cite{Paulavicius2011,Evtushenko2009,Jones2009}) constructing the essentially nonuniform coverages.
+This holds even in the case of using the efficient algorithms (for example, \cite{Paulavicius2011,Evtushenko2009,Jones2009,Paulavicius2020,Sergeyev2021}) constructing the essentially nonuniform coverages.
 
@@
 
-In the present study, we used the global search algorithm combined with the dimensionality reduction scheme based on Peano curves \cite{Sergeyev2013,Lera2021} to solve the global optimization problem (\ref{global_problem}).
-This algorithm was demonstrated to be a powerful tool for solving complex multiextremal problems \cite{Kvasov2013,Kalyulin2017,Cavoretto2021}.
+In the present study, we used the global search algorithm combined with the dimensionality reduction scheme based on Peano curves \cite{Sergeyev2013,Lera2015,Lera2021} to solve the global optimization problem (\ref{global_problem}).
+This algorithm was demonstrated to be a powerful tool for solving complex multiextremal problems (see \cite{Kvasov2013,Gillard2016,Kalyulin2017,Cavoretto2021}).
 To solve the local problem (\ref{local_problem}), we used Hooke-Jeeves (pattern search) method (see, e.g., \cite{Kelley}). 
@@
 The choice of the point is performed by one of the following methods:
-\begin{enumerate}
+\begin{itemize}
 \item the point is fixed according to the physical meaning of the applied problem being solved;
@@
 \item use the solution found by the global method (e.g. global search algorithm) in $K$ iterations as the reference point.
-\end{enumerate}
+\end{itemize}
 
@@
 \begin{enumerate}
-\item compute $P+1$ values of the objective function $\varphi(y)$ at the points $z_i^j = (\overline{y}_1,...,\overline{y}_{i-1},y_i^j,\overline{y}_{i+1},...,\overline{y}_N)$, where \mbox{$y_i^j = a_i + jh$},  $h=(b_i-a_i)/P, \; 0\leq j \leq P$;
-\item form a set $Q$ comprising of $P+1$ pairs of the kind $\left\{y_i^j, \varphi(z_i^j)\right\} $ 
-\item construct a regression model of given power $deg$ based on the data from $Q$;
-\item compute the coefficient of determination $R^2$ and perform the classification of the $i^{\rm th}$ variable:
+\item[2.1] compute $P+1$ values of the objective function $\varphi(y)$ at the points $z_i^j = (\overline{y}_1,...,\overline{y}_{i-1},y_i^j,\overline{y}_{i+1},...,\overline{y}_N)$, where \mbox{$y_i^j = a_i + jh$},  $h=(b_i-a_i)/P, \; 0\leq j \leq P$;
+\item[2.2] form a set $Q$ comprising of $P+1$ pairs of the kind $\left\{y_i^j, \varphi(z_i^j)\right\} $ 
+\item[2.3] construct a regression model of given power $deg$ based on the data from $Q$;
+\item[2.4] compute the coefficient of determination $R^2$ and perform the classification of the $i^{\rm th}$ variable:
 \begin{itemize}
@@
 
-\begin{table}[ht]
+\begin{table}
 	\caption{Results of solving the series of test problems}

### 2022-07-21 2022/ICNAAM/ICNAAM.tex
after reviewing

@@
 % Title portion
-\title{Using Machine Learning for Separation of Parameters in Global Optimization Problems of Large Dimensionality}
+\title{Using Machine Learning for Separation of Parameters in High-Dimensional Global Optimization Problems}
 
@@
 \begin{abstract}
-The work presents the results of investigations of an approach to solving the optimization problems, in which the dependencies on different groups of parameters have different nature.
-A scheme for selecting the parameters of the problem, which have local effect on the objective function, has been proposed.
-It allows to solve the essentially multidimensional problems using the nested optimization scheme.
-At that, different optimization algorithms are used at different nesting levels.
+This paper examines the findings from research into an approach to solving optimization problems, in which the dependencies on diverse groups of parameters are of a different nature.
+A scheme has been proposed for choosing the problem parameters, which have a local effect on the objective function.
+It allows solving essentially  multidimensional problems using the nested optimization scheme.
+In this case,  different optimization algorithms are used at differing nesting levels.
 The proposed approach has demonstrated its efficiency in solving several series of test problems.
@@
 
-At present, the global optimization methods are applied to solving various problems arising in various fields of science and technology (see \cite{Pinter2006}). 
+At present, global optimization methods are applied to solving a variety of problems arising in various fields of science and technology (see \cite{Pinter2006}). 
 For example, the evaluation of the model parameters from the experimental data is a traditional field of application of the global optimization methods.
-Among these ones, there are the problems of chemical kinetics (see \cite{Gubaydullin2021}).
-In the problems of this kind, it is necessary to perform a search for the values of unknown model parameters, at which the modeling results are the closest to 
-the ones obtained experimentally.
+Such applications include chemical kinetics problems (see \cite{Gubaydullin2021}).
+In the problems of this kind, it is necessary to perform a search for the values of unknown model parameters, at which the modeling results are closest to those obtained experimentally.
 
-The number of parameters, which are to be evaluated in this way, can be tens and hundreds for the chemical kinetics problems.
-The use of the deterministic global optimization methods for solving the problems of such a dimensionality is very limited because of very large computation costs of covering the search domain by the trial points.
-This holds even in the case of using the efficient algorithms (for example, \cite{Paulavicius2011,Evtushenko2009,Jones2009,Paulavicius2020,Sergeyev2021}) constructing the essentially nonuniform coverages.
+The number of parameters to be evaluated in this way can be tens and hundreds for chemical kinetics problems.
+The use of deterministic global optimization methods for solving the problems of such a dimensionality is very limited because of very large computation costs of covering the search domain by trial points.
+This holds even in the case when efficient algorithms (for example, \cite{Paulavicius2011,Evtushenko2009,Jones2009,Paulavicius2020,Sergeyev2021}) are used to construct the essentially nonuniform coverages.
 
-Numerous inverse problems are featured by different dependencies on different groups of parameters. 
-For example, the dependence on one group of parameters may be close to the linear one whereas the dependence on another group of parameters may have a complex multiextremal character.
+Numerous inverse problems are characterized by different dependencies on varying groups of parameters. 
+For example, the dependence on one group of parameters may be close to the linear one whereas the dependence on another group of parameters may have a complex multiextremal nature.
+In this case, the problem can be solved using the recursive optimization scheme. 
+A multiextremal subproblem will be solved at the upper nesting level with the use of complex global optimization algorithms.
+Unimodal subproblems (each corresponding to a fixed set of the values of the first part of parameters) will be solved at the lower nesting level.
+To solve such problems, one can apply local optimization methods having a fast convergence.
 
-In this case, the problem can be solved using the recursive optimization scheme. 
-The solving of the multiextremal subproblem, which requires the use of complex global optimization algorithms, will be performed at the upper nesting level.
-The unimodal subproblems (each corresponding to a fixed set of the values of the first part of parameters) will be solve at the lower nesting level.
-To solve these ones, one can apply the local optimization methods having a fast convergence.
-
-As a rule, one can't subdivide the parameters into the groups with different character of the objective function behavior in advance since the objective function in applied problems is usually defined as a black box.
-So far, the development of the algorithm for the minimization of the essentially multiextremal functions taking into account different character of the dependence of the objective function on different groups of parameters is relevant.
+As a rule, one cannot subdivide the parameters into groups with different nature of the objective function behavior in advance since the objective function in applied problems is usually defined as a black box.
+To date, the development of an algorithm for the minimization of essentially multiextremal functions taking into account the different nature of the dependence of the objective function on various groups of parameters is quite relevant.
 In the present work, a special method for such a separation based on the analysis of the regression model of the function with respect to each variable separately has been proposed.
@@
 with the constant $L$ unknown {\em a priori}.
-At the same time, let us assume the dependence of the objective function on $N_1$ parameters to be multiextremal while the dependence on the rest $N_2 = N - N_1$ parameters is close to the linear or quadratic one.
+At the same time, let us assume the dependence of the objective function on $N_1$ parameters to be multiextremal while the dependence on the remaining $N_2 = N - N_1$ parameters is close to the linear or quadratic one.
 At the same time, the subdivision of the parameters is unknown. 
@@
 Taking into account such a specific feature of the problem can essentially reduce the computation costs of the search for the optimum. 
-Indeed, according to well known nested optimization scheme \cite{Grishagin2001,Grishagin2015}, the solving of the initial problem can be performed in the following way:
+Indeed, according to well known nested optimization scheme \cite{Grishagin2001,Grishagin2015}, the initial problem can be solved in the following way:
 \begin{equation}\label{global_problem}
@@
 %(this process will be called {\em a search trial})
-implies the solving of a unimodal problem (\ref{local_problem}), which can be performed by one of local optimization methods.
-The efficiency of the local optimization methods allow solving the subproblems (\ref{local_problem}) with a precision essentially exceeding the one of the global optimization methods.
-Accordingly, we can consider the problem (\ref{global_problem}) as the global optimization one where the process of computing an objective function value to be a 
+implies solving the unimodal problem (\ref{local_problem}), which can be performed by one of local optimization methods.
+The efficiency of local optimization methods makes it possible to solve the subproblems (\ref{local_problem}) with an accuracy essentially exceeding the accuracy of global optimization methods.
+Accordingly, we can consider the problem (\ref{global_problem}) as a global optimization problem where the process of computing an objective function value is a 
 time-consuming operation.
@@
 This algorithm was demonstrated to be a powerful tool for solving complex multiextremal problems (see \cite{Kvasov2013,Gillard2016,Kalyulin2017,Cavoretto2021}).
-To solve the local problem (\ref{local_problem}), we used Hooke-Jeeves (pattern search) method (see, e.g., \cite{Kelley}). 
-Unlike the BFGS-type methods (see, e.g., \cite{Nocedal}), this algorithm doesn't require computing the gradient and works well under conditions of the 
-computing errors in the function values that is typical in solving the applied problems.
+To solve the local problem (\ref{local_problem}), we used the Hooke-Jeeves (pattern search) method (see, e.g., \cite{Kelley}). 
+Unlike the BFGS-type methods (see, e.g., \cite{Nocedal}), this algorithm does not require computing the gradient and works well under the conditions of computational errors in the function values, which is typical in solving applied problems.
 
-\section{SEPARATION OF THE PROBLEMS PARAMETERS}
+\section{SEPARATION OF THE PROBLEM PARAMETERS}
 
@@
 \item the point is fixed according to the physical meaning of the applied problem being solved;
-\item use the solution found by the local method (e.g. Hooke-Jeeves method) in $K$ iterations as the reference point;
-\item use the solution found by the global method (e.g. global search algorithm) in $K$ iterations as the reference point.
+\item use the solution found by the local method (e.g. the  Hooke-Jeeves method) in $K$ iterations as the reference point;
+\item use the solution found by the global method (e.g. the global search algorithm) in $K$ iterations as the reference point.
 \end{itemize}
@@
 \item[2.1] compute $P+1$ values of the objective function $\varphi(y)$ at the points $z_i^j = (\overline{y}_1,...,\overline{y}_{i-1},y_i^j,\overline{y}_{i+1},...,\overline{y}_N)$, where \mbox{$y_i^j = a_i + jh$},  $h=(b_i-a_i)/P, \; 0\leq j \leq P$;
-\item[2.2] form a set $Q$ comprising of $P+1$ pairs of the kind $\left\{y_i^j, \varphi(z_i^j)\right\} $ 
-\item[2.3] construct a regression model of given power $deg$ based on the data from $Q$;
+\item[2.2] form a set $Q$ comprising $P+1$ pairs of the kind $\left\{y_i^j, \varphi(z_i^j)\right\} $ 
+\item[2.3] construct a regression model of a given power $deg$ based on the data from $Q$;
 \item[2.4] compute the coefficient of determination $R^2$ and perform the classification of the $i^{\rm th}$ variable:
@@
 \item if $R^2$ is greater than a predefined threshold $T$, then add the variable to the set of the local variables;
-\item else add the variable to the set of the global variables.
+\item else add the variable to the set of global variables.
 \end{itemize}
@@
 If the local variables are found, then continue solving the problem using the nested optimization scheme; the global search algorithm will be used at the upper level while the local optimization method will be used at the lower one.
-Otherwise, run the global search algorithm taking all variables as the global ones.
+Otherwise, run the global search algorithm, assuming all variables are global.
 
@@
 The quality of the model constructed was evaluated at Step 2, Item 4 using the adjusted coefficient of determination $R^2_a$.
-This coefficient features the similarity of the regression model constructed to the objective function in the fixed section.
+This coefficient characterizes the similarity of the regression model constructed to the objective function in the fixed section.
 The value of $R^2_a$, as a rule, falls between 0 and 1.
@@
 The values of $K$, $deg$, and $T$ are the parameters of the method and are set by the researcher according to the properties of the problem being solved.
-The algorithm allows using the power of the regression polynomial $deg = 1$ or $deg = 2$ that allows constructing the model based on either linear or quadratic regression, respectively.
+The algorithm makes it possible to use the power of the regression polynomial $deg = 1$ or $deg = 2$, which permits the construction of a model based on either linear or quadratic regression, respectively.
 
@@
 
-In the first set of the test problems (we will denote it as Gr-L), we used 100 20-dimensional combined-type problems: the local subproblems of the dimensionality $N_2 = 18$ and a global subproblem of the dimensionality $N_1=2$.
+In the first set of the test problems (we will denote it as Gr-L), we used 100 20-dimensional combined-type problems: a local subproblem of the dimensionality $N_2 = 18$ and a global subproblem of the dimensionality $N_1=2$.
 The 18-dimensional subproblems comprising the combinations of nearly linear functions were used as the local component:
@@
 The multiextremal functions proposed by Grishagin (see, for example, \cite{Grishagin1994}) were used as the global components of the problems $\varphi_j^{glob}(y)$. 
-
 The objective function $\varphi(y)$ comprised a linear combination of these subproblems, i.e.
 $\varphi_j(y) = \alpha \varphi_j^{glob}(y) + \beta \varphi_j^{loc}(y)$.
-The sequence of the parameters in each problem was selected randomly.
+The sequence of parameters in each problem was selected randomly.
+Level lines of a multiextremal subproblem and a two-dimensional section of a local subproblem are shown in Figure \ref{fig}.
 
-In the second set of problems (hereafter called Gr-LQ), the parameters $M_{ij}$ and $L_{ij}$ of the local subproblem (\ref{X2_problem}) were organized so that in 60\% of the cases the viable had an insufficient quadratic part, in 30\% of the cases the contributions of the quadratic components were essential (at least, comparable to the contribution of the linear parts), and the rest 10\% of variables were purely linear.
+\begin{figure}
+\begin{minipage}{0.27\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1a.png}}
+\end{minipage}
 
-In the third series of experiments (hereafter called GKLS-L), 4-dimensional functions from the GKLS class (see, for example, in \cite{Sergeyev2015,Grishagin2018}) were used as the multiextremal parts.
-As the local components, the local subproblems of the Gr-L series were used.
+\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;
+
+\begin{minipage}{0.27\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1b.png}}
+\end{minipage}
+\caption{Level lines of a multiextremal subproblem and a two-dimensional section of a local subproblem.}
+\label{fig}
+\end{figure}
+
+In the second set of problems (hereafter called Gr-LQ), the parameters $M_{ij}$ and $L_{ij}$ of the local subproblem (\ref{X2_problem}) were organized so that in 60\% of the cases the variable had an insufficient quadratic part, in 30\% of the cases the contributions of the quadratic components were essential (at least, comparable to the contribution of the linear parts), and the remaining 10\% of variables were purely linear.
+
+In the third series of experiments (hereafter called GKLS-L), 4-dimensional functions from the GKLS class (see, for example, \cite{Sergeyev2015,Grishagin2018}) were used as the multiextremal parts.
+The local subproblems of the Gr-L series were used as local components.
 The series included 20 problems.
 
-As a result of solving the Gr-L series with the preset precision of $0.01$ and the parameters of the method $K=100$, $deg=1$, and $T=0.6$, correct separation of the variables into the global and local ones were performed in all problems.
-Then, all problems were solved successfully using the nested optimization scheme in appropriate time and number of iterations.
-However, the solving of the complicated series Gr-LQ with the use of the linear regression didn't give good results because of the presence of the parabolic-type variables, for which the percentages of similarity to the linear model were extremely low.
+As a result of solving the Gr-L series with the preset accuracy of $0.01$ and the parameters of the method $K=100$, $deg=1$, and $T=0.6$, correct separation of the variables into the global and local ones was performed in all problems.
+Then, all problems were solved successfully using the nested optimization scheme in appropriate time and with the appropriate number of iterations.
+However, solving the complicated series Gr-LQ with the use of the linear regression did not yield good results because of the presence of the parabolic-type variables, for which the percentages of similarity to the linear model were extremely low.
 In this connection, other parameters of the method $K=100$, $deg=2$ (the quadratic regression), and $T=0.95$ were selected to conduct the experiments.
@@
 
-In the third series of experiments, the whole series of the problems was not solved successfully in appropriate time and number of iterations because of the specific features of the construction of the GKLS functions.
-Most variables of the GKLS functions in different sections are strongly similar to the quadratic functions. 
-Therefore, one has to raise the threshold $T$ to exclude an accidental discarding a global variable.
+In the third series of experiments, the whole series of the problems could not be solved successfully in the 	appropriate time and with an appropriate number of iterations because of the specific features of the construction of the GKLS functions.
+Most variables of the GKLS functions in different sections are strongly similar to quadratic functions. 
+Therefore, one has to raise the threshold $T$ to exclude an accidental discarding of a global variable.
 As a consequence, some local parameters were classified as the global ones as well.
-A large number of global variables contribute to the increase in the number of iterations and in the problem solving time naturally.
-On the other hand, if one sets the threshold for including a variable into the local functions low enough, the method wouldn't find the global variables in some functions at all.
+Obviously, a large number of global variables contributes to the increase in the number of iterations and in the problem solving time.
+On the other hand, if one sets the threshold for including a variable into the local functions low enough, the method would not find the global variables in some functions at all.
 
 The numerical results of the series of experiments are presented in Table \ref{tab1}.
-In this one, the data on the number of problems solved $S$, the averaged number of iterations of the global search at the upper nesting level $G_{av}$, the averaged number of trials at the lower nesting level $L_{av}$, and the averaged time of solving a single problem $t_{av}$ (in seconds) are presented.
-The computations were carried out using a computer with CPU Intel Core i7 10750H, 2.6 GHz. 
-The scikit-learn package from Python 3.9 was used for the constructing and analysis of the regression model. 
-The global search algorithm and nested optimization scheme were implemented using C++.
+In this table, the data on the number of problems solved $S$, the averaged number of iterations of the global search at the upper nesting level $G_{av}$, the averaged number of trials at the lower nesting level $L_{av}$, and the averaged time of solving a single problem $t_{av}$ (in seconds) are presented.
+The computations were carried out using a computer with an Intel Core i7 10750H CPU at 2.6 GHz. 
+The scikit-learn package from Python 3.9 was used to construct and analyze the regression model. 
+The global search algorithm and the nested optimization scheme were implemented using C++.
 

### 2022-07-26 2022/ICNAAM_NetOpt/article/Надумин.tex
первая версия статьи

% !!!!!!!!!!!!!!!!!!   ВНИМАНИЕ !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
% Заголовки разделов формируются при помощи команд \section{}, \subsection{}, \subsubsection{}
% Не используйте уровень вложенности заголовков больше трех!
% -----------------------------
% Для оформления теорем, лемм, следствий используйте окружения 
% Def     - Определение
% Teor    - Теорема
% Lem     - Лемма
% Predl   - Предложение
% Ass     - Утверждение
% Cor     - Следствие
% Example - Пример
% -----------------------------
% Доказательство теоремы начинается командой \proof и завершается командой \endproof
% -----------------------------
% Литература помещается в окружение biblio.

\documentclass[11pt, oneside, a4paper]{article}
%\usepackage[cp1251]{inputenc} % кодировка
\usepackage[utf8]{inputenc} % кодировка
\usepackage[english, russian]{babel} % Русские и английские переносы
\usepackage{graphicx}          % для включения графических изображений
\usepackage{cite}              % для корректного оформления литературы
\usepackage{enumitem}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{mathtext}
\usepackage{array}
\usepackage{hyperref}

%стилевой пакет
\usepackage{schoolseminar2022}

\begin{document}
% \udk     - универсальный десятичный классификатор
% \msc     -
% \title   - название статьи
% \authors - список авторов

\setcounter{page}{1}

\udk{519.853.4, 004.032.26}

\title{Использование аппроксимации нейронными сетями при решении задач глобальной оптимизации\footnote{Работа выполнена при поддержке Министерства науки и высшего образования РФ (проект \textnumero~0729-2020-0055) и научно-образовательного математического центра <<Математика технологий будущего>> (проект \textnumero~075-02-2021-1394).}}

\authors{Карпенко С.Н., Лебедев И.Г., Надумин Д.В.}
\organizations{ННГУ им. Н.И. Лобачевского\\
Институт информационных технологий, математики и механики}

\abstractru{
В работе рассматривается задача поиска глобального минимума многоэкстремальной функции. Для решения этой задачи был выбран алгоритм имитации отжига. Для ускорения процесса поиска минимума используется аппроксимация целевой функции нейросетью. Представлены результаты работы комбинированного алгоритма глобальной оптимизации, объединяющего алгоритм имитации отжига, использующего нейронную сеть для аппроксимации функций и метод Нелдера-Мида для точного решения задачи оптимизации.  В данной работе проводились эксперименты с серией двумерных функций из генератора GKLS. Результаты показывают сокращение числа вычислений целевой функции при сохранении высокой точности поиска.
}

\keywords{методы отпимизации, глобальная оптимизация, нейронные сети, машинное обучение.}

% \section{название} - заголовок раздела первого уровня
% \subsection{название} - заголовок раздела второго уровня
% \subsubsection{название} - заголовок раздела третьего уровня
% Не используйте уровень вложенности заголовков больше трех!
% Каждый абзац текста в статье начинается командой \par или пустой
% строкой.

\bigskip

\section{Введение}

Нейронная сеть представляет собой математическую модель, построенную по принципу организации и функционирования биологических нейронных сетей – сетей нервных клеток живого организма. Это понятие возникло при изучении процессов, протекающих в мозге, и при попытке смоделировать эти процессы.

С точки зрения математики, обучение нейронных сетей - это многопараметрическая задача нелинейной оптимизации. С точки зрения развития вычислительной техники и программирования, нейронная сеть - способ решения проблемы эффективного параллелизма.

Нейронные сети применяются для решения многих прикладных задач. Наиболее известными сферами применений нейронных сетей являются: распознавание образов и классификация, прогнозирование, аппроксимация.
Нейронные сети могут аппроксимировать непрерывные функции. С помощью линейных операций и каскадного соединения можно из произвольного нелинейного элемента получить устройство, вычисляющее любую непрерывную функцию с некоторой наперёд заданной точностью.

История нейронных сетей намного длиннее, чем принято считать. Сама идея "способной к мышлению системы" возникла еще в Древней Греции, и популярность нейронных сетей менялась с течением времени.

В 1943 году Уоррен Маккалок и Уолтер Питтс опубликовали работу «Логическое исчисление идей, относящихся к нервной деятельности» \cite{fio_bib3}. Целью данного исследования было изучение работы человеческого мозга. В 1958 году, Фрэнк Розенблатт в своем исследовании «Персептрон: вероятностная модель хранения и организации информации в головном мозге» \cite{fio_bib4} описал модель персептрона. В 1974 году, первым ученым на территории США, описавшим в своей диссертации \cite{fio_bib5} использование алгоритма обратного распространения ошибки в нейронных сетях, стал Пол Вербос, хотя развитием этой идеи занимались многие исследователи. В 1989 году, Янн Лекун опубликовал статью \cite{fio_bib6}, в которой было описано практическое использование ограничений обратного распространения ошибки и интеграция в архитектуру нейронной сети для обучения алгоритмов.

\section{Постановка задачи глобальной оптимизации}
Задача многомерной многоэкстремальной оптимизации может быть определена как проблема поиска наименьшего значения действительной функции $\phi(y)$

\begin{equation*}   %%%%%%%%
\phi(y^*) = min\{\phi(y):y\in D\},
\end{equation*}
\begin{equation}   %%%%%%%%
D = \{y \in R^N : a_i \leq y_i \leq b_i,
1 \leq i \leq N \},
\end{equation}

где $a,b \in R$ - заданные векторы.

Численное решение задачи (1) к построению оценки $y_k^* \in D$, отвечающей некоторому понятию близости к точке $y^*$ (например, $\parallel y^*-y_k^* \parallel \leq \epsilon$, где $\epsilon>0$ - некоторая заданная точность) на основе конечного числа $k$ вычислений оптимизируемой функции. Относительно класса рассматриваемых задач предполагается выполнение двух важных условий:
\begin{enumerate}
\item Предполагается, что оптимизируемая функция $\phi(y)$ может быть задана не аналитически, в виде формулы, а алгоритмически, как результат работы некоторой подпрограммы или библиотеки.

\item Будем предполагать, что $\phi(y)$ удовлетворяет условию Липшица
\begin{equation}
|\phi(y_1)-\phi(y_2)|\leq L\parallel y_1-y_2 \parallel
,y_1,y_2 \in D, 0<L< \infty
\end{equation}
\end{enumerate}
что соответствует ограниченности изменения значений функции при ограниченной вариации аргумента. Это предположение можно интерпретировать (применительно к прикладным задачам) как отражение ограниченности мощностей, порождающих изменения в моделируемой системе.

Задачи многоэкстремальной оптимизации имеют существенно более высокую трудоемкость решения по сравнению с другими типами оптимизационных задач, т.к. глобальный оптимум является интегральной характеристикой решаемой задачи и требует исследования всей области поиска. Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (сетки) в области параметров, и выборе наилучшего значения функции на данной сетке. Вычислительные затраты на решение задачи растут экспоненциально с ростом размерности \cite{fio_bib7}.

\section{Методы локальной и глобальной оптмизиации}
\subsection{Оптимизация методом имитации отжига}
Алгоритм имитации отжига является общим алгоритмическим методом решения задачи глобальной оптимизации.

Алгоритм основывается на имитации физического процесса, который происходит при кристаллизации вещества, в том числе при отжиге металлов.

При помощи моделирования такого процесса ищется такая точка или множество точек, на котором достигается минимум некоторой числовой функции $f(x^i),x^i=(x_1,...x_n)\in X$. Решение ищется последовательным вычислением точек пространства $X$\cite{fio_bib16}.

Введём следующие обозначения: $X$ - множество вссех состояний системы, $x_i$-состояние на $i$-ом шаге алгоритма; $t_i$-температура на $i$-ом шаге.

Определим три функции:
\begin{enumerate}
\item Функция энергии(то, что оптимизируем):   $E:X\rightarrow R$
\item Функция изменения температуры с течением времени:  
$T:N\rightarrow R$
\item Функция,порождающая новое состояние:  
$F:X\rightarrow X$
\end{enumerate}
На входе минимальная $t_{min} $ и начальная $t_{max}$ температуры. Задаём произвольное первое состояние $x_1$ и $t_{min}=t_{max}$.

Пока $t_i>t_{min}$:
\begin{enumerate}
\item $x_c=F(x_{i-1})$,
\item $\triangle E=E(x_c)-E(x_{i-1})$,
\item Если $\triangle E \leq 0$, то $x_i=x_c$. Если $\triangle E >0$, то переход осуществляется с вероятностью $P(\triangle E)=e^{- \triangle E /t_i}$,
\item Понижаем температуру $t_{i+1}=T(i)$.
\end{enumerate}
Возвращаем последнее состояние $x$. 
\subsection{Метод Нелдера-Мида}
Метод Нелдера-Мида является методом безусловной оптимизации \cite{fio_bib19}функции нескольких  переменных $f(x^1,...,x^n )$, не использующий градиентов функции. Суть метода заключается в последовательном перемещении и деформировании симплекса вокруг точки экстремума. Метод находит локальный экстремум и может «застрять» в одном из них. Предполагается, что серьёзных ограничений на область определения функции нет, то есть функция определена во всех встречающихся точках.
Параметрами метода являются:
\begin{enumerate}
\item Коэффициент отражения $\alpha$>0, обычно выбирается равным 1.
\item Коэффициент сжатия $\beta$>0, обычно выбирается равным 0,5.
\item Коэффициент растяжения $\gamma$>0, обычно выбирается равным 2.
\end{enumerate}
\subsection{Комбинированный алгоритм глобальной оптимизации}
Нами разработан комбинированный алгоритм глобальной оптимизации, использующий нейронную сеть для аппроксимации функций, метод имитации отжига для поиска глобального оптимума и метод Нелдера-Мида для уточнения найденного решения.

В данной работе была выбрана нейронная сеть прямого распространения (feedforward neural network)\cite{fio_bib2,fio_bib3,fio_bib4,fio_bib5,fio_bib6,fio_bib7} из библиотеки Keras\cite{fio_bib15}.

В ходе исследования параметры нейронной сети были подобраны следующим образом: 5 слоёв по 32 нейрона в каждом слое. В качестве активационной функции выбрана сигмоида. Весовые коэффициенты имеют нормальное распределение со средним равным 0 и стандартным отклонением равным 0,2, что позволяет задавать их значения от -1 до 1. Смещение нейрона (bias) инициализируется 0. В выходном слое только один нейрон.

Сам комбинированный алгоритм можно разбить на три пункта:
\begin{enumerate}
\item Создание обучающей выборки для нейронной сети.
\item Обучение нейронной сети.
\item Пока не выполнен критерий остановки, либо пока не выполнено ограничение на количество вычислений целевой функции делаем следующее:
\begin{enumerate}
\item Оптимизация нейросетевой аппроксимации с помощью алгоритма имитации отжига.
\item Оптимизация целевой функции с помощью локального метода, где в качестве первоначального значения используется найденная глобальным методом точка.
\item Если разница между найденным значением и глобальным минимумом меньше 0,01, то комбинированный алгоритм сошёлся.
\item Иначе добавляем в обучающую выборку 10 случайных точек и найденную локальным методом точку.
\item После чего, заново обучаем нейронную сеть и уходим на следующую итераци.
\end{enumerate}
\end{enumerate}
\section{Описание программной реализации}
Программная реализация выполнена на языке Python с помощью библиотек Numpy \cite{fio_bib12}, Matplotlib \cite{fio_bib13}, SciPy \cite{fio_bib14}, Keras \cite{fio_bib15}.

Работа выполнялась в среде Jupyter Notebook. Программа разбита на 5 блоков:
\begin{enumerate}
\item Импортируются все библиотеки, в том числе пользовательские функции и генератор задач GKLS.
\item Определена функция задания нейросети.
\item Задаётся обучающая выборка.
\item Задаются параметры нейросети и критерии остановки комбинированного алгоритма.
\item Выполнение комбинированного алгоритма.
\end{enumerate}

\section{Эксперименты}
Эксперименты проводились на компьютере со следующими характеристиками:
\begin{enumerate}
\item Процессор Intel Core i3-9100F, 4100 MHz.
\item ОЗУ 16 Gb (2x8 Gb) DDR4.
\item Видеокарта GeForce GTX 1050 2 Gb.
\end{enumerate}

Задача экспериментов состояла в том, чтобы уменьшить количество вычислений целевой функции.

На рис.1 представлена поверхность одной из функции генератора GKLS, на рис.2 - её аппроксимация нейросетью.

\begin{figure}[!h]
	\begin{center}
		\begin{minipage}[!h]{0.45\linewidth}
			\includegraphics[width=1\linewidth]{figure/1.png}
			\caption{Поверхность функции.} %% подпись к рисунку
		\end{minipage}
		\hfill
		\begin{minipage}[!h]{0.45\linewidth}
			\includegraphics[width=1\linewidth]{figure/1_approx.png}
			\caption{Аппроксимация.}
		\end{minipage}
	\end{center}
\end{figure}

Эксперименты проводились на 10-ти двумерных задачах из генератора GKLS.
\begin{table}
	\caption{Результаты работы имитации отжига и комбинированного алгоритма}
	\begin{center}
		\begin{tabular}{|m{3.8cm}|m{6cm} |m{6cm}|}
			\hline
			 & Среднее количество вычислений целевой функции & Точность нахождения решений\\
			 \hline
			 Имитация отжига & 4255,9 & $1,8 \cdot 10^{-8}$\\
			 \hline 
			 Комбинированный алгоритм & 293,4 & $3,3 \cdot 10^{-5}$\\
			 \hline
		\end{tabular}
	\end{center}
\end{table}
Результаты экспериментов показывают значительное уменьшение количества вычислений целевой функции. При этом сохраняется высокая точность нахождения минимума. В дальнейшем планируются эксперименты с детерминированным алгоритмом глобального поиска вместо вероятностного алгоритма имитации отжига.

\begin{biblio}
\bibitem{fio_bib1}
Мордухович Б. Ш. Методы аппроксимаций в задачах оптимизации и управления / Мордухович Б. Ш. . – М.: «Наука», 1988. – 360 c.

\bibitem{fio_bib2}
Cybenko, G. V. Approximation by Superpositions of a Sigmoidal function. Mathematics of Control Signals and Systems. — 1989. — Т. 2, № 4. — С. 303—314.

\bibitem{fio_bib3}
Warren S. McCulloch, Walter Pitts // California State University. – Режим доступа: \url{https://home.csulb.edu/~cwallis/382/readings/482/mccolloch.logical.calculus.ideas.1943.pdf}

\bibitem{fio_bib4}
F. Rosenblatt // Cornell Aeronautical Laboratory. – Режим доступа: \url{https://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.335.3398&rep=rep1type=pdf}

\bibitem{fio_bib5}
Paul Werbos // National Science Foundation. – Режим доступа: \url{https://www.researchgate.net/publication/35657389_Beyond_regression_new_tools_for_prediction_and_analysis_in_the_behavioral_sciences}

Y. LeCun, B. Boser // AT \& T  Bell Laboratories. – Режим доступа: \url{http://yann.lecun.com/exdb/publis/pdf/lecun-89e.pdf}

\bibitem{fio_bib7}
Гермейер Ю.Б. Введение в теорию исследования операций / Гермейер Ю.Б. – М.: «Наука», 1971. – 384 c.

\bibitem{fio_bib8}
Сухарев А. Г. Оптимальный поиск экстремума / Сухарев. А. Г. Изд. МГУ, М, 1975.

\bibitem{fio_bib9}
Стронгин Р. Г. Численные методы в многоэкстремальных задачах / Стронгин Р. Г. – М.: «Наука»,1978. – 240 с.

\bibitem{fio_bib10}
Информационный подход к задаче поиска экстремума функций, Изв. АН СССР, Техническая кибернетка, 1 (1996), 17-26.

\bibitem{fio_bib11}
К.А. Баркалов, И.Г. Лебедев. Решение задач глобальной оптимизации 
на графических ускорителях – ННГУ им. Н.И. Лобачевского.

\bibitem{fio_bib12}
Официальный сайт библиотеки NumPy. – Режим доступа: \url{https://numpy.org/}

\bibitem{fio_bib13}
Официальный сайт библиотеки Matplotlib. – Режим доступа: \url{https://matplotlib.org/}

\bibitem{fio_bib14}
Официальный сайт библиотеки SciPy. – Режим доступа: \url{https://scipy.org/}

\bibitem{fio_bib15}
Официальный сайт библиотеки Keras. – Режим доступа: \url{https://keras.io/}

\bibitem{fio_bib16}
Kirkpatrick, S.; Gelatt Jr,C.D.;Vecchi, M. P. (1983). "Optimization by Simulated Annealing". Science. 220 (4598): 671–680.

\bibitem{fio_bib17}
Cергеев Ю.В., Квасов Д.Е. Диагональные  методы глобальной оптимизиации. Физматлит, 2008. P. 352.

\bibitem{fio_bib18}
Gaviano M., Lera D., Kvasov D. E., Sergeyev Y. D. Software for generation of classes of test functions with known local and global minima for global optimization // ACM Transactions on Mathematical Software. 2003. Vol. 29. P. 469-480.

\bibitem{fio_bib19}
J. A. Nelder and R. Mead, Computer Journal, 1965, vol. 7, p. 308—313.

\end{biblio}

\newpage

\msc{90C26, 82C32}

\title{The using neural network approximation in global optimization problems}

\authors{S.N.~Karpenko,  I.G.~Lebedev, D.V.~Nadumin}
\organizations{N. I. Lobachevsky State University of Nizhny Novgorod\\ 	Institute of information technologies, mathematics and mechanics}

% abstract is contained in the  abstract environment
\abstracten{
The paper considers the problem of finding the global minimum of a multi-extremal function. To speed up the process of finding the minimum, the approximation of the objective function by a neural network is used.
The article presents the results of a combined global optimization algorithm using a neural network to approximate functions.

The works \cite{fio_bib17,fio_bib18} describe a GKLS generator that allows generating multi-extreme optimization problems with pre-known properties: the number of local minima, the size of their areas of attraction, the point of the global minimum, the value of the function in it, etc. In this paper, experiments were carried out with a series of two-dimensional functions from the GKLS generator

The work was performed in the Python programming language with Numpy, SciPy, Keras, Matplotlib libraries.
The purpose of the experiments was to test the possibility of reducing the number of calculations of the objective function when solving the optimization problem.}
\keywordsen{optimization methods, global optimization, neural networks, machine learning.}

\begin{biblioen}

\bibitem{fio_bib1}
Mordukhovich B. S. Approximation methods in optimization and control problems / Mordukhovich B. S. - M.: "Science", 1988. – 360 p.

\bibitem{fio_bib2}
Cybenko, G. V. Approximation by Superpositions of a Sigmoidal function. Mathematics of Control Signals and Systems. — 1989. — Т. 2, № 4. — С. 303—314.

\bibitem{fio_bib3}
Warren S. McCulloch, Walter Pitts // California State University. – Режим доступа: \url{https://home.csulb.edu/~cwallis/382/readings/482/mccolloch.logical.calculus.ideas.1943.pdf}

\bibitem{fio_bib4}
F. Rosenblatt // Cornell Aeronautical Laboratory. – Режим доступа: \url{https://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.335.3398&rep=rep1type=pdf}

\bibitem{fio_bib5}
Paul Werbos // National Science Foundation. – Режим доступа: \url{https://www.researchgate.net/publication/35657389_Beyond_regression_new_tools_for_prediction_and_analysis_in_the_behavioral_sciences}

\bibitem{fio_bib6}
Y. LeCun, B. Boser // AT \& T  Bell Laboratories. – Режим доступа: \url{http://yann.lecun.com/exdb/publis/pdf/lecun-89e.pdf}

\bibitem{fio_bib7}
Hermeyer Yu.B. Introduction to the theory of operations research / Hermeyer Yu.B. – M.: "Science", 1971. – 384 p.

\bibitem{fio_bib8}
Sukharev A. G. Optimal search for an extremum / Sukharev. A. G. Publishing House of Moscow State University, Moscow, 1975.

\bibitem{fio_bib9}
Strongin R. G. Numerical methods in multi–extreme problems / Strongin R. G. – M.: "Science", 1978. - 240 p.

\bibitem{fio_bib10}
Information approach to the problem of finding the extremum of functions, Izv. of the USSR Academy of Sciences, Technical Cybernetics, 1 (1996), 17-26.

\bibitem{fio_bib11}
K.A. Barkalov, I.G. Lebedev. Solving global optimization
problems on graphics accelerators – N.I. Lobachevsky National Research University.

\bibitem{fio_bib12}
The official website of the NumPy library. – Access mode: \url{https://numpy.org /}

\bibitem{fio_bib13}
The official website of the Matplotlib library. – Access mode: \url{https://matplotlib.org /}

\bibitem{fio_bib14}
The official website of the SciPy library. – Access mode: \url{https://scipy.org /}

\bibitem{fio_bib15}
The official website of the Keras Library. – Access mode: \url{https://keras.io /}

\bibitem{fio_bib16}
Kirkpatrick, S.; Gelatt Jr,C.D.;Vecchi, M. P. (1983). "Optimization by Simulated Annealing". Science. 220 (4598): 671–680.

\bibitem{fio_bib17}
Sergeev Yu.V., Kvasov D.E. Diagonal methods of global optimization. Fizmatlit, 2008. P. 352.

\bibitem{fio_bib18}
Gaviano M., Lera D., Kvasov D. E., Sergeyev Y. D. Software for generation of classes of test functions with known local and global minima for global optimization // ACM Transactions on Mathematical Software. 2003. Vol. 29. P. 469-480.

\bibitem{fio_bib19}
J. A. Nelder and R. Mead, Computer Journal, 1965, vol. 7, p. 308—313.

\end{biblioen}

\end{document}

### 2022-07-29 2022/ICNAAM_NetOpt/article/SchoolseminarNetOpt.tex
Обновленная версия статьи.

--- 2022-07-26 2022/ICNAAM_NetOpt/article/Надумин.tex
@@
 
-\keywords{методы отпимизации, глобальная оптимизация, нейронные сети, машинное обучение.}
+\keywords{методы оптимизации, глобальная оптимизация, нейронные сети, машинное обучение.}
 
@@
 \end{equation}
-
 где $a,b \in R$ - заданные векторы.
 
-Численное решение задачи (1) к построению оценки $y_k^* \in D$, отвечающей некоторому понятию близости к точке $y^*$ (например, $\parallel y^*-y_k^* \parallel \leq \epsilon$, где $\epsilon>0$ - некоторая заданная точность) на основе конечного числа $k$ вычислений оптимизируемой функции.
+Численное решение задачи (1) к построению оценки $y_k^* \in D$, отвечающей некоторому понятию близости к точке $y^*$ (например, $\parallel y^*-y_k^* \parallel \leq \epsilon$, где $\epsilon>0$ это некоторая заданная точность) на основе конечного числа $k$ вычислений оптимизируемой функции.
 Относительно класса рассматриваемых задач предполагается выполнение двух важных условий:
 \begin{enumerate}
-\item Предполагается, что оптимизируемая функция $\phi(y)$ может быть задана не аналитически, в виде формулы, а алгоритмически, как результат работы некоторой подпрограммы или библиотеки.
+\item Предполагается, что оптимизируемая функция $\phi(y)$ может быть задана алгоритмически, как результат работы некоторой подпрограммы или библиотеки.
 
@@
 
-\section{Методы локальной и глобальной оптмизиации}
-\subsection{Оптимизация методом имитации отжига}
+\section{Метод имитации отжига}
 Алгоритм имитации отжига является общим алгоритмическим методом решения задачи глобальной оптимизации.
+Он основывается на имитации физического процесса, который происходит при кристаллизации вещества, в том числе при отжиге металлов.
 
-Алгоритм основывается на имитации физического процесса, который происходит при кристаллизации вещества, в том числе при отжиге металлов.
+При помощи моделирования такого процесса ищется такая точка или множество точек, на котором достигается минимум некоторой числовой функции $\phi(y^i),y^i=(y_1,...y_n)\in Y$.
+Решение ищется последовательным вычислением точек пространства $Y$\cite{fio_bib16}, \cite{fio_bib20}.
 
-При помощи моделирования такого процесса ищется такая точка или множество точек, на котором достигается минимум некоторой числовой функции $f(x^i),x^i=(x_1,...x_n)\in X$.
-Решение ищется последовательным вычислением точек пространства $X$\cite{fio_bib16}.
-
-Введём следующие обозначения: $X$ - множество вссех состояний системы, $x_i$-состояние на $i$-ом шаге алгоритма; $t_i$-температура на $i$-ом шаге.
+Введём следующие обозначения: $Y$ - множество всех состояний системы, $y_i$-состояние на $i$-ом шаге алгоритма; $t_i$-температура на $i$-ом шаге.
 
@@
 \begin{enumerate}
-\item Функция энергии(то, что оптимизируем):   $E:X\rightarrow R$
+\item Функция энергии(то, что оптимизируем):   $E:Y\rightarrow R$
 \item Функция изменения температуры с течением времени:  
 $T:N\rightarrow R$
-\item Функция,порождающая новое состояние:  
-$F:X\rightarrow X$
+\item Функция, порождающая новое состояние:  
+$\phi:Y\rightarrow Y$
 \end{enumerate}
-На входе минимальная $t_{min} $ и начальная $t_{max}$ температуры. Задаём произвольное первое состояние $x_1$ и $t_{min}=t_{max}$.
+На входе минимальная $t_{min} $ и начальная $t_{max}$ температуры. Задаём произвольное первое состояние $y_1$ и $t_{min}=t_{max}$.
 
@@
 \begin{enumerate}
-\item $x_c=F(x_{i-1})$,
-\item $\triangle E=E(x_c)-E(x_{i-1})$,
-\item Если $\triangle E \leq 0$, то $x_i=x_c$. Если $\triangle E >0$, то переход осуществляется с вероятностью $P(\triangle E)=e^{- \triangle E /t_i}$,
+\item $y_c=\phi(y_{i-1})$,
+\item $\triangle E=E(y_c)-E(y_{i-1})$,
+\item Если $\triangle E \leq 0$, то $y_i=y_c$. Если $\triangle E >0$, то переход осуществляется с вероятностью $P(\triangle E)=e^{- \triangle E /t_i}$,
 \item Понижаем температуру $t_{i+1}=T(i)$.
 \end{enumerate}
-Возвращаем последнее состояние $x$. 
-\subsection{Метод Нелдера-Мида}
-Метод Нелдера-Мида является методом безусловной оптимизации \cite{fio_bib19}функции нескольких  переменных $f(x^1,...,x^n )$, не использующий градиентов функции.
-Суть метода заключается в последовательном перемещении и деформировании симплекса вокруг точки экстремума.
-Метод находит локальный экстремум и может «застрять» в одном из них.
-Предполагается, что серьёзных ограничений на область определения функции нет, то есть функция определена во всех встречающихся точках.
-Параметрами метода являются:
-\begin{enumerate}
-\item Коэффициент отражения $\alpha$>0, обычно выбирается равным 1.
-\item Коэффициент сжатия $\beta$>0, обычно выбирается равным 0,5.
-\item Коэффициент растяжения $\gamma$>0, обычно выбирается равным 2.
-\end{enumerate}
-\subsection{Комбинированный алгоритм глобальной оптимизации}
+Возвращаем последнее состояние $y$.
+
+Преимуществом метода имитации отжига является возможность избежать локальные минимумы оптимизируемой функции за счет принятия изменений, временно ухудшающих результат, что отражает суть процесса нагрева расплава для предотвращения его быстрого остывания.
+Еще одним преимущество – даже в условиях нехватки вычислительных ресурсов для нахождения глобального минимума метод отжига выдает один из локальных минимумов.
+
+\section{Метод Нелдера-Мида}
+Метод Нелдера-Мида является методом безусловной оптимизации \cite{fio_bib19} функции нескольких  переменных $\phi(y^1,...,y^n )$, не использующий градиентов функции.
+Он относится к группе методов прямого поиска, так как в нем не используются производные.
+Метод является развитием симплексного метода.
+
+Идея метода заключается в следующем: в пространстве $R^N$ выбираются $(n+1)$ пробных векторов, в каждом из которых можно вычислить значение целевой функции.
+Вершина с наихудшим значение целевой функции -- $A$ отражается через центр тяжести других вершин многогранника, в результате чего из оставшихся вершин строится новый многогранник, называемый отраженным, и одна точка, расположенная на линии проекции на соответствующем расстоянии от центра тяжести.
+
+Метод находит локальный экстремум и может "застрять" в одном из них. Предполагается что функция определена во всех точках области определения.
+
+\section{Математический аппарат нейронной сети прямого распространения}
+Согласно универсальной теореме аппроксимации (теореме Цыбенко), доказанной Джорджем Цыбенко в 1989 году, искусственная нейронная сеть с прямым распространением может аппроксимировать любую непрерывную функцию многих переменных с любой точностью.
+Условиями являются: достаточное количество нейронов скрытого слоя, удачный подбор весов и смещений.
+
+На входе нейронная сеть принимает набор значений переменной.
+Далее эти значения умножаются на некоторые весовые коэффициенты.
+После чего в качестве взвешенной по весовым коэффициентам суммы подаются  на вход  каждого из нейронов скрытого слоя, где к этой сумме прибавляется смещение (bias).
+В результате получаем число, которое подаётся на вход активационной функции.
+Результат активационной функции и есть выходные значение одного нейрона.
+
+Взвешенная сумма представлена ниже:
+\begin{equation}
+z = \sum_{i=1}^n y_i \cdot w_i +bias,
+\end{equation}
+где $bias$ – смещение, $w_i$ – весовой коэффициент, $y_i$ - входное значение.
+
+Далее (3) подаётся в качестве входного значения в активационную функцию.
+Существует несколько видов активационных функций.
+В данной работе применялась сигмоидальная функции, область значений которой от 0 до 1.
+\begin{equation}
+\sigma(z)= \frac{1}{1+e^{-z}}
+\end{equation}
+
+\section{Комбинированный алгоритм глобальной оптимизации}
 Нами разработан комбинированный алгоритм глобальной оптимизации, использующий нейронную сеть для аппроксимации функций, метод имитации отжига для поиска глобального оптимума и метод Нелдера-Мида для уточнения найденного решения.
@@
 \item Иначе добавляем в обучающую выборку 10 случайных точек и найденную локальным методом точку.
-\item После чего, заново обучаем нейронную сеть и уходим на следующую итераци.
+\item После чего, заново обучаем нейронную сеть и уходим на следующую итерации.
 \end{enumerate}
@@
 \begin{enumerate}
-\item Импортируются все библиотеки, в том числе пользовательские функции и генератор задач GKLS.
-\item Определена функция задания нейросети.
-\item Задаётся обучающая выборка.
+\item Импортируются все библиотеки, в том числе пользовательские функции (в частности, функция нормализации значений) и генератор задач GKLS. 
+\item Определена функция задания нейросети, принимающая в качестве аргумента размерность задачи. Функция возвращает объект нейронной сети.
+\item Задаётся обучающая выборка, состоящая из 225 значений в области определения функций генератора GKLS.
 \item Задаются параметры нейросети и критерии остановки комбинированного алгоритма.
-\item Выполнение комбинированного алгоритма.
+Количество эпох обучения равно 6000.
+Размер выборки для обучения (batch size) на каждой эпохе (каждой итерации) равен одной трети от всего размера обучающей выборки.
+\item Выполнение комбинированного алгоритма. В случае повторения обучения нейросети batch size также задаётся равным одной трети от всей выборки с учётом добавленных точек.
 \end{enumerate}
@@
 На рис.1 представлена поверхность одной из функции генератора GKLS, на рис.2 - её аппроксимация нейросетью.
-
 \begin{figure}[!h]
 	\begin{center}
-		\begin{minipage}[!h]{0.45\linewidth}
+		\begin{minipage}[h]{0.45\linewidth}
 			\includegraphics[width=1\linewidth]{figure/1.png}
@@
 
+\newpage
+
 Эксперименты проводились на 10-ти двумерных задачах из генератора GKLS.
+В работе \cite{fio_bib21} \cite{fio_bib22} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.
+
+В таблице 1 отражено среднее число вычислений целевой функции, которые выполнили метод имитации отжига и разработанный нами комбинированный алгоритм, при решении серии задач.
+Также приведена средняя точность нахождения решения.
+Максимальное число итераций алгоритма имитации отжига составляло $10000$, остальные параметры соответствуют значениям по умолчанию.
+
 \begin{table}
@@
 \end{table}
-Результаты экспериментов показывают значительное уменьшение количества вычислений целевой функции.
-При этом сохраняется высокая точность нахождения минимума.
-В дальнейшем планируются эксперименты с детерминированным алгоритмом глобального поиска вместо вероятностного алгоритма имитации отжига.
+
+Результаты экспериментов показывают значительное уменьшение количества вычислений целевой функции. При этом сохраняется высокая точность нахождения минимума.
 
@@
 
+\bibitem{fio_bib20}
+Simon Haykin. A Comprehensive Foundation (2nd ed.). Prentice Hall PTR, Upper
+Saddle River, Neural Networks, NJ, USA, 1998
+
+\bibitem{fio_bib21}
+Gaviano, M.
+Software for generation of classes of test functions with known local and global minima for global optimization/ M.
+Gaviano, D.
+Lera, D.
+E.
+Kvasov, Y.
+D.
+Sergeyev // ACM Transactions on Mathematical Software.
+– 2003.
+– Vol.
+29.
+– P.
+469-480.
+
+\bibitem{fio_bib22}
+Сергеев, Я.Д. Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. – 352 c.
+
 \end{biblio}
@@
 \abstracten{
-The paper considers the problem of finding the global minimum of a multi-extremal function.
+The paper considers the problem of finding the global minimum of a multiextremal function.
+To solve this problem, an annealing simulation algorithm was chosen.
 To speed up the process of finding the minimum, the approximation of the objective function by a neural network is used.
-The article presents the results of a combined global optimization algorithm using a neural network to approximate functions.
-
-The works \cite{fio_bib17,fio_bib18} describe a GKLS generator that allows generating multi-extreme optimization problems with pre-known properties: the number of local minima, the size of their areas of attraction, the point of the global minimum, the value of the function in it, etc.
-In this paper, experiments were carried out with a series of two-dimensional functions from the GKLS generator
-
-The work was performed in the Python programming language with Numpy, SciPy, Keras, Matplotlib libraries.
-The purpose of the experiments was to test the possibility of reducing the number of calculations of the objective function when solving the optimization problem.}
+The results of a combined global optimization algorithm combining an annealing simulation algorithm using a neural network to approximate functions and the Nelder-Meade method for the exact solution of the optimization problem are presented.
+In this paper, experiments were carried out with a series of two-dimensional functions from the GKLS generator.
+The results show a reduction in the number of calculations of the objective function while maintaining high search accuracy.}
 \keywordsen{optimization methods, global optimization, neural networks, machine learning.}
@@
 
+\bibitem{fio_bib20}
+Simon Haykin. A Comprehensive Foundation (2nd ed.). Prentice Hall PTR, Upper
+Saddle River, Neural Networks, NJ, USA, 1998
+
+\bibitem{fio_bib21}
+Gaviano, M.
+Software for generation of classes of test functions with known local and global minima for global optimization/ M.
+Gaviano, D.
+Lera, D.
+E.
+Kvasov, Y.
+D.
+Sergeyev // ACM Transactions on Mathematical Software.
+– 2003.
+– Vol.
+29.
+– P.
+469-480.
+
+\bibitem{fio_bib22}
+Сергеев, Я.Д. Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. – 352 c.
+
 \end{biblioen}

### 2022-07-31 2022/ICNAAM_NetOpt/article/SchoolseminarNetOpt.tex
Поправлен список литературы

@@
 Эксперименты проводились на 10-ти двумерных задачах из генератора GKLS.
-В работе \cite{fio_bib21} \cite{fio_bib22} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.
+В работе \cite{fio_bib19} \cite{fio_bib20} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.
 
@@
 \begin{biblio}
-\bibitem{fio_bib1}
-Мордухович Б. Ш. Методы аппроксимаций в задачах оптимизации и управления / Мордухович Б. Ш. . – М.: «Наука», 1988. – 360 c.
 
@@
 
+\bibitem{fio_bib6}
 Y. LeCun, B. Boser // AT \& T  Bell Laboratories. – Режим доступа: \url{http://yann.lecun.com/exdb/publis/pdf/lecun-89e.pdf}
@@
 Гермейер Ю.Б. Введение в теорию исследования операций / Гермейер Ю.Б. – М.: «Наука», 1971. – 384 c.
-
-\bibitem{fio_bib8}
-Сухарев А. Г. Оптимальный поиск экстремума / Сухарев. А. Г. Изд. МГУ, М, 1975.
-
-\bibitem{fio_bib9}
-Стронгин Р. Г. Численные методы в многоэкстремальных задачах / Стронгин Р. Г. – М.: «Наука»,1978. – 240 с.
-
-\bibitem{fio_bib10}
-Информационный подход к задаче поиска экстремума функций, Изв. АН СССР, Техническая кибернетка, 1 (1996), 17-26.
-
-\bibitem{fio_bib11}
-К.А. Баркалов, И.Г. Лебедев. Решение задач глобальной оптимизации 
-на графических ускорителях – ННГУ им. Н.И. Лобачевского.
 
@@
 
-\bibitem{fio_bib17}
-Cергеев Ю.В., Квасов Д.Е. Диагональные  методы глобальной оптимизиации. Физматлит, 2008. P. 352.
-
-\bibitem{fio_bib18}
-Gaviano M., Lera D., Kvasov D.
-E., Sergeyev Y.
-D.
-Software for generation of classes of test functions with known local and global minima for global optimization // ACM Transactions on Mathematical Software.
-2003.
-Vol.
-29.
-P.
-469-480.
-
 \bibitem{fio_bib19}
-J. A. Nelder and R. Mead, Computer Journal, 1965, vol. 7, p. 308—313.
-
-\bibitem{fio_bib20}
-Simon Haykin. A Comprehensive Foundation (2nd ed.). Prentice Hall PTR, Upper
-Saddle River, Neural Networks, NJ, USA, 1998
-
-\bibitem{fio_bib21}
 Gaviano, M.
@@
 
-\bibitem{fio_bib22}
+\bibitem{fio_bib20}
 Сергеев, Я.Д. Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. – 352 c.
@@
 
-\bibitem{fio_bib1}
-Mordukhovich B. S. Approximation methods in optimization and control problems / Mordukhovich B. S. - M.: "Science", 1988. – 360 p.
-
 \bibitem{fio_bib2}
@@
 
-\bibitem{fio_bib8}
-Sukharev A. G. Optimal search for an extremum / Sukharev. A. G. Publishing House of Moscow State University, Moscow, 1975.
-
-\bibitem{fio_bib9}
-Strongin R. G. Numerical methods in multi–extreme problems / Strongin R. G. – M.: "Science", 1978. - 240 p.
-
-\bibitem{fio_bib10}
-Information approach to the problem of finding the extremum of functions, Izv. of the USSR Academy of Sciences, Technical Cybernetics, 1 (1996), 17-26.
-
-\bibitem{fio_bib11}
-K.A. Barkalov, I.G. Lebedev. Solving global optimization
-problems on graphics accelerators – N.I. Lobachevsky National Research University.
-
 \bibitem{fio_bib12}
@@
 
-\bibitem{fio_bib17}
-Sergeev Yu.V., Kvasov D.E. Diagonal methods of global optimization. Fizmatlit, 2008. P. 352.
-
-\bibitem{fio_bib18}
-Gaviano M., Lera D., Kvasov D.
-E., Sergeyev Y.
-D.
-Software for generation of classes of test functions with known local and global minima for global optimization // ACM Transactions on Mathematical Software.
-2003.
-Vol.
-29.
-P.
-469-480.
-
 \bibitem{fio_bib19}
-J. A. Nelder and R. Mead, Computer Journal, 1965, vol. 7, p. 308—313.
-
-\bibitem{fio_bib20}
-Simon Haykin. A Comprehensive Foundation (2nd ed.). Prentice Hall PTR, Upper
-Saddle River, Neural Networks, NJ, USA, 1998
-
-\bibitem{fio_bib21}
 Gaviano, M.
@@
 
-\bibitem{fio_bib22}
-Сергеев, Я.Д. Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. – 352 c.
+\bibitem{fio_bib20}
+Sergeev Yu.V., Kvasov D.E. Diagonal methods of global optimization. Fizmatlit, 2008. P. 352.
 

### 2022-08-02 2022/ICNAAM_NetOpt/article/SchoolseminarNetOpt.tex
fix some errors

@@
 
-\title{Использование аппроксимации нейронными сетями при решении задач глобальной оптимизации\footnote{Работа выполнена при поддержке Министерства науки и высшего образования РФ (проект \textnumero~0729-2020-0055) и научно-образовательного математического центра <<Математика технологий будущего>> (проект \textnumero~075-02-2021-1394).}}
+\title{Использование нейросетевой аппроксимации целевой функции при решении задач глобальной оптимизации\footnote{Работа выполнена при поддержке Министерства науки и высшего образования РФ (проект \textnumero~0729-2020-0055) и научно-образовательного математического центра <<Математика технологий будущего>> (проект \textnumero~075-02-2021-1394).}}
 
@@
 
-Нейронная сеть представляет собой математическую модель, построенную по принципу организации и функционирования биологических нейронных сетей – сетей нервных клеток живого организма.
+Нейронная сеть представляет собой математическую модель, построенную по принципу организации и функционирования биологических нейронных сетей --- сетей нервных клеток живого организма.
 Это понятие возникло при изучении процессов, протекающих в мозге, и при попытке смоделировать эти процессы.
 
-С точки зрения математики, обучение нейронных сетей - это многопараметрическая задача нелинейной оптимизации.
-С точки зрения развития вычислительной техники и программирования, нейронная сеть - способ решения проблемы эффективного параллелизма.
+С точки зрения математики, обучение нейронных сетей --- это многопараметрическая задача нелинейной оптимизации.
+С точки зрения развития вычислительной техники и программирования, нейронная сеть --- способ решения проблемы эффективного параллелизма.
 
 Нейронные сети применяются для решения многих прикладных задач. Наиболее известными сферами применений нейронных сетей являются: распознавание образов и классификация, прогнозирование, аппроксимация.
-Нейронные сети могут аппроксимировать непрерывные функции.
-С помощью линейных операций и каскадного соединения можно из произвольного нелинейного элемента получить устройство, вычисляющее любую непрерывную функцию с некоторой наперёд заданной точностью.
+Нейронные сети могут аппроксимировать непрерывные функции: с помощью линейных операций и каскадного соединения можно из произвольного нелинейного элемента получить устройство, вычисляющее любую непрерывную функцию с некоторой наперёд заданной точностью.
 
-История нейронных сетей намного длиннее, чем принято считать. Сама идея "способной к мышлению системы" возникла еще в Древней Греции, и популярность нейронных сетей менялась с течением времени.
+История нейронных сетей намного длиннее, чем принято считать. Сама идея <<способной к мышлению системы>> возникла еще в Древней Греции, и популярность нейронных сетей менялась с течением времени.
 
-В 1943 году Уоррен Маккалок и Уолтер Питтс опубликовали работу «Логическое исчисление идей, относящихся к нервной деятельности» \cite{fio_bib3}.
+В 1943 году Уоррен Маккалок и Уолтер Питтс опубликовали работу <<Логическое исчисление идей, относящихся к нервной деятельности>> \cite{fio_bib3}.
 Целью данного исследования было изучение работы человеческого мозга.
@@
 Задача многомерной многоэкстремальной оптимизации может быть определена как проблема поиска наименьшего значения действительной функции $\phi(y)$
-
 \begin{equation*}   %%%%%%%%
@@
 \end{equation}
-где $a,b \in R$ - заданные векторы.
+где $a,b \in R$ --- заданные векторы.
 
@@
 \begin{enumerate}
-\item Предполагается, что оптимизируемая функция $\phi(y)$ может быть задана алгоритмически, как результат работы некоторой подпрограммы или библиотеки.
+\item целевая функция $\phi(y)$ может быть задана алгоритмически, как результат работы некоторой подпрограммы или библиотеки;
 
-\item Будем предполагать, что $\phi(y)$ удовлетворяет условию Липшица
+\item целевая функция $\phi(y)$ удовлетворяет условию Липшица
 \begin{equation}
 |\phi(y_1)-\phi(y_2)|\leq L\parallel y_1-y_2 \parallel
-,y_1,y_2 \in D, 0<L< \infty
+,y_1,y_2 \in D, 0<L< \infty.
 \end{equation}
 \end{enumerate}
-что соответствует ограниченности изменения значений функции при ограниченной вариации аргумента.
+Второе предположение соответствует ограниченности изменения значений функции при ограниченной вариации аргумента.
 Это предположение можно интерпретировать (применительно к прикладным задачам) как отражение ограниченности мощностей, порождающих изменения в моделируемой системе.
@@
 \section{Метод имитации отжига}
+
+%эту фразу надо исправить
 Алгоритм имитации отжига является общим алгоритмическим методом решения задачи глобальной оптимизации.
@@
 
-Введём следующие обозначения: $Y$ - множество всех состояний системы, $y_i$-состояние на $i$-ом шаге алгоритма; $t_i$-температура на $i$-ом шаге.
+Введём следующие обозначения: $Y$ --- множество всех состояний системы, $y_i$ --- состояние на $i$-ом шаге алгоритма; $t_i$ --- температура на $i$-ом шаге.
 
@@
 \begin{enumerate}
-\item Функция энергии(то, что оптимизируем):   $E:Y\rightarrow R$
+\item Функция энергии(то, что оптимизируем):   $E:Y\rightarrow R$.
 \item Функция изменения температуры с течением времени:  
-$T:N\rightarrow R$
+$T:N\rightarrow R$.
 \item Функция, порождающая новое состояние:  
-$\phi:Y\rightarrow Y$
+$\phi:Y\rightarrow Y$.
 \end{enumerate}
-На входе минимальная $t_{min} $ и начальная $t_{max}$ температуры. Задаём произвольное первое состояние $y_1$ и $t_{min}=t_{max}$.
+На входе задаются минимальная $t_{min} $ и начальная $t_{max}$ температуры. Задаём произвольное первое состояние $y_1$ и $t_{min}=t_{max}$.
 
@@
 
-Преимуществом метода имитации отжига является возможность избежать локальные минимумы оптимизируемой функции за счет принятия изменений, временно ухудшающих результат, что отражает суть процесса нагрева расплава для предотвращения его быстрого остывания.
-Еще одним преимущество – даже в условиях нехватки вычислительных ресурсов для нахождения глобального минимума метод отжига выдает один из локальных минимумов.
+Особенностью метода имитации отжига является возможность выхода из локального минимума оптимизируемой функции за счет принятия изменений, временно ухудшающих результат, что отражает суть процесса нагрева расплава для предотвращения его быстрого остывания.
+Еще одно преимущество --- даже в условиях нехватки вычислительных ресурсов для нахождения глобального минимума метод отжига выдает один из локальных минимумов.
 
 \section{Метод Нелдера-Мида}
+
 Метод Нелдера-Мида является методом безусловной оптимизации \cite{fio_bib19} функции нескольких  переменных $\phi(y^1,...,y^n )$, не использующий градиентов функции.
@@
 Идея метода заключается в следующем: в пространстве $R^N$ выбираются $(n+1)$ пробных векторов, в каждом из которых можно вычислить значение целевой функции.
-Вершина с наихудшим значение целевой функции -- $A$ отражается через центр тяжести других вершин многогранника, в результате чего из оставшихся вершин строится новый многогранник, называемый отраженным, и одна точка, расположенная на линии проекции на соответствующем расстоянии от центра тяжести.
+Вершина $A$ с наихудшим значение целевой функции отражается через центр тяжести других вершин многогранника, в результате чего из оставшихся вершин строится новый многогранник, называемый отраженным, и одна точка, расположенная на линии проекции на соответствующем расстоянии от центра тяжести.
+При этом предполагается что функция определена во всех точках области определения.
 
-Метод находит локальный экстремум и может "застрять" в одном из них. Предполагается что функция определена во всех точках области определения.
+Метод Нелдера-Мида относится к методам локальной оптимизации, т.е. он сходится, вообще говоря, к локальному экстремуму функции.
 
 \section{Математический аппарат нейронной сети прямого распространения}
+
 Согласно универсальной теореме аппроксимации (теореме Цыбенко), доказанной Джорджем Цыбенко в 1989 году, искусственная нейронная сеть с прямым распространением может аппроксимировать любую непрерывную функцию многих переменных с любой точностью.
@@
 \begin{equation}
-z = \sum_{i=1}^n y_i \cdot w_i +bias,
+z = \sum_{i=1}^n y_i \cdot w_i + bias,
 \end{equation}
-где $bias$ – смещение, $w_i$ – весовой коэффициент, $y_i$ - входное значение.
+где $bias$ --- смещение, $w_i$ --- весовой коэффициент, $y_i$ --- входное значение.
 
@@
 \section{Комбинированный алгоритм глобальной оптимизации}
+
 Нами разработан комбинированный алгоритм глобальной оптимизации, использующий нейронную сеть для аппроксимации функций, метод имитации отжига для поиска глобального оптимума и метод Нелдера-Мида для уточнения найденного решения.
@@
 В качестве активационной функции выбрана сигмоида.
-Весовые коэффициенты имеют нормальное распределение со средним равным 0 и стандартным отклонением равным 0,2, что позволяет задавать их значения от -1 до 1.
+Весовые коэффициенты имеют нормальное распределение со средним равным 0 и стандартным отклонением равным 0.2, что позволяет задавать их значения в диапазоне $[-1,1]$.
 Смещение нейрона (bias) инициализируется 0.
@@
 
-На рис.1 представлена поверхность одной из функции генератора GKLS, на рис.2 - её аппроксимация нейросетью.
+На рис.~1 представлена поверхность одной из функций, порожденных генератором GKLS, а на рис.~2 --- её аппроксимация нейросетью.
 \begin{figure}[!h]
@@
 Эксперименты проводились на 10-ти двумерных задачах из генератора GKLS.
-В работе \cite{fio_bib19} \cite{fio_bib20} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.
+В работах \cite{fio_bib19}, \cite{fio_bib20} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.
 
@@
 			 \hline
-			 Имитация отжига & 4255,9 & $1,8 \cdot 10^{-8}$\\
+			 Имитация отжига & 4255.9 & $1.8 \cdot 10^{-8}$\\
 			 \hline 
-			 Комбинированный алгоритм & 293,4 & $3,3 \cdot 10^{-5}$\\
+			 Комбинированный алгоритм & 293.4 & $3.3 \cdot 10^{-5}$\\
 			 \hline
@@
 
-\title{The using neural network approximation in global optimization problems}
+\title{Using neural network approximation of the objective function in global optimization problems}
 
@@
 The paper considers the problem of finding the global minimum of a multiextremal function.
-To solve this problem, an annealing simulation algorithm was chosen.
-To speed up the process of finding the minimum, the approximation of the objective function by a neural network is used.
-The results of a combined global optimization algorithm combining an annealing simulation algorithm using a neural network to approximate functions and the Nelder-Meade method for the exact solution of the optimization problem are presented.
-In this paper, experiments were carried out with a series of two-dimensional functions from the GKLS generator.
+To solve this problem, the simulation annealing  algorithm was used.
+To speed up the process of finding the minimum, the approximation of the objective function by a neural network is applied.
+The results of algorithm that combines the simulation annealing  algorithm using a neural network to approximate objective function and the Nelder-Meade method to refine the solution of the optimization problem are presented.
+The paper presentS the results of the experiments carried out with a series of multiextremal test problems.
 The results show a reduction in the number of calculations of the objective function while maintaining high search accuracy.}

### 2022-08-04 2022/Saransk/full_paper_Barkalov_Usova/paper.tex
add first version of full paper

% !!!!!!!!!!!!!!!!!!   ВНИМАНИЕ !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
% Заголовки разделов формируются при помощи команд \section{}, \subsection{}, \subsubsection{}
% Не используйте уровень вложенности заголовков больше трех!
% -----------------------------
% Для оформления теорем, лемм, следствий используйте окружения 
% Def     - Определение
% Teor    - Теорема
% Lem     - Лемма
% Predl   - Предложение
% Ass     - Утверждение
% Cor     - Следствие
% Example - Пример
% -----------------------------
% Доказательство теоремы начинается командой \proof и завершается командой \endproof
% -----------------------------
% Литература помещается в окружение biblio.

\documentclass[11pt, oneside, a4paper]{article}
%\usepackage[cp1251]{inputenc} % кодировка
\usepackage[utf8]{inputenc} % кодировка
\usepackage[english, russian]{babel} % Русские и английские переносы
\usepackage{graphicx}          % для включения графических изображений
\usepackage{cite}              % для корректного оформления литературы
\usepackage{enumitem}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{mathtext}

%стилевой пакет
\usepackage{schoolseminar2022}

\begin{document}
% \udk     - универсальный десятичный классификатор
% \msc     -
% \title   - название статьи
% \authors - список авторов

\setcounter{page}{1}

\title{Разделение параметров в задачах глобальной оптимизации с помощью методов машинного обучения
\renewcommand{\thefootnote}{*}\footnote{Работа выполнена при поддержке Министерства науки и высшего образования РФ (проект \textnumero~0729-2020-0055) и научно-образовательного математического центра <<Математика технологий будущего>> (проект \textnumero~075-02-2021-1394).}}

\authors{К.А.~Баркалов, М.А.~Усова}
\organizations{Нижегородский государственный университет им. Н.И. Лобачевского \\ Нижний Новгород, Россия}

\abstractru{
В работе представлены результаты исследования подхода к решению задач глобальной оптимизации с разным характером зависимости от разных групп параметров.
Предложена схема выделения параметров задачи, которые оказывают локальное влияние на целевую функцию, что позволяет решать существенно многомерные задачи с использованием многошаговой схемы редукции размерности. При этом на разных уровнях рекурсии используются разные оптимизационные алгоритмы. Проведено исследование работоспособности предложенного подхода.
}

\keywords{глобальная оптимизация, локальная оптимизация, редукция размерности, разделение параметров}

% \section{название} - заголовок раздела первого уровня
% \subsection{название} - заголовок раздела второго уровня
% \subsubsection{название} - заголовок раздела третьего уровня
% Не используйте уровень вложенности заголовков больше трех!
% Каждый абзац текста в статье начинается командой \par или пустой
% строкой.

\bigskip

\section{Постановка задачи}

В настоящее время методы глобальной оптимизации используются в различных областях науки и техники, например, для идентификации значений параметров математических моделей, при которых результаты моделирования наиболее близки к результатам, полученным экспериментально.
Число параметров, которые требуется определить подобным образом, например, для задач химической кинетики, может составлять десятки и сотни. В данном случае использование детерминированных методов глобальной оптимизации крайне ограничено в силу чрезвычайно больших вычислительных затрат на покрытие области поиска точками испытаний. Это остается справедливым даже в случае использования эффективных алгоритмов (например, \cite{Evtushenko2009,Paulavicius2016}), строящих существенно неравномерные покрытия. 
При этом частым явлением в подобных задачах является близкая к линейной зависимость по некоторой группе параметров, в то время как для остальных параметров наблюдается сложный многоэкстремальный характер. Заранее указать разделение на группы параметров с разным характером поведения целевой функции, как правило, не представляется возможным, т.к. целевая функция в обратных задачах задается как черный ящик.

Рассмотрим задачу оптимизации вида
\begin{eqnarray}\label{main_problem}
& \varphi(y^\ast)=\min{\left\{\varphi(y): y\in D\right\}}, \nonumber \\
& D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
\end{eqnarray}
Будем предполагать, что функция $\varphi(y)$ является многоэкстремальной, заданной в виде черного ящика и удовлетворяет условию Липшица
\[
\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in D,\; 0<L<\infty,
\]
с априори неизвестной константой $L$.
Целевые функции большинства практических задач оптимизации обладают данным свойством.
Одновременно с этим будем предполагать, что зависимость целевой функции от $N_1$ параметров является многоэкстремальной, а от оставшихся $N_2 = N - N_1$ параметров зависимость является близкой к линейной или квадратичной. При этом разделение  параметров является неизвестным.

\section{Общее описание подхода к решению}

Решение поставленной задачи можно организовать с помощью многошаговой схемы редукции размерности \cite{Grishagin2007}.

Решение  многоэкстремальной подзадачи, для которой требуется использовать сложные алгоритмы глобальной оптимизации, будет проводиться на верхнем уровне рекурсии с помощью алгоритма глобального поиска \cite{Strongin13} в сочетании со схемой редукции размерности на основе кривых Пеано \cite{Sergeyev2013}. Данный алгоритм зарекомендовал себя как мощный инструмент для решения сложных многоэкстремальных задач.

Унимодальные подзадачи будут решаться на нижнем уровне с помощью метода Хука-Дживса. В отличие от методов типа BFGS, данный алгоритм не требует вычисления градиента и хорошо работает в условиях наличия вычислительных ошибок в значениях функций, что является типичным при решении прикладных задач.

Данный подход поможет существенно снизить вычислительную сложность поиска оптимума.

\section{Разделение параметров задачи}

Для разделения параметров задачи на локальные и глобальные предлагается следующая последовательность действий:

\textbf{Этап 1.} Зафиксировать опорную точку $\overline{y} = (\overline{y}_1, \overline{y}_2,...,\overline{y}_N)$ для целевой функции $\varphi(y)$ одним из следующих способов:
\begin{itemize}
\item точка фиксируется исходя из физического смысла решаемой прикладной задачи;
\item в качестве фиксированной точки используется решение, найденное локальным методом за $K$ итераций;
\item в качестве фиксированной точки используется решение, найденное глобальным методом за $K$ итераций.
\end{itemize}

\textbf{Этап 2.} Для всех компонент (переменных) $y_i, \; 1\leq i \leq N$ провести исследование на локальность:
\begin{enumerate}
\item вычислить $P+1$ значение целевой функции $\varphi(y)$ в точках $z_i^j = (\overline{y}_1,...,\overline{y}_{i-1},y_i^j,\overline{y}_{i+1},...,\overline{y}_N)$, где
$y_i^j =  a_i + jh, \; h=(b_i-a_i)/P, \; 0\leq j \leq P$;
\item сформировать множество $Q$, представляющее собой набор из $P+1$ пары вида $\left\{y_i^j, \varphi(z_i^j)\right\} $;
\item построить регрессионную модель заданной степени $deg$ на основе данных из $Q$;
\item вычислить оценку $R^2$ и провести классификацию i-й переменной:
\begin{itemize}
\item если $R^2$ больше заданного порога $T$, то добавить переменную к множеству локальных переменных,
\item иначе добавить переменную к множеству глобальных переменных.
\end{itemize}
\end{enumerate}

Если локальные переменные были обнаружены, то продолжить решение задачи с использованием описанной ранее многошаговой схемы. В противном случае запустить алгоритм глобального поиска, считая все переменные глобальными.

Значения  $K$, $deg$ и $T$ являются параметрами метода и устанавливаются исследователем в соответствии со свойствами решаемой задачей. В реализованном методе $deg = 1$ или $deg = 2$, что позволяет построить модель линейной или квадратичной регрессии соответственно.

\begin{figure}[h!]
\centering \includegraphics[width=1\linewidth]{regression_94}
\caption{Линейные регрессионные модели 15-мерной тестовой задачи  (2 глобальных параметра)}\label{regression_94_}
\end{figure}
\begin{figure}[h!]
\centering \includegraphics[width=1\linewidth]{regression_94_3}
\caption{Квадратичные регрессионные модели 15-мерной тестовой задачи (2 глобальных параметра)}\label{regression_94_3_}
\end{figure}

В основе проводимой регрессии лежит метод наименьших квадратов (МНК). Оценка качества построенной на этапе 2 модели производится с помощью скорректированного коэффициента детерминации $R^2_a$. Указанный коэффициент характеризует схожесть построенной модели регрессии с целевой функцией в зафиксированном сечении.

\section{Численные эксперименты}

Предложенный подход показал свою работоспособность при решении нескольких серий тестовых задач, представляющих собой линейные комбинации подзадач с глобальными параметрами (функции Гришагина или GRIS, функции Сергеева или GKLS \cite{Grishagin2001}) и подзадач с локальными параметрами, представляющих собой комбинацию близких к линейным одномерных функций (далее L) или  комбинацию  функций с разным вкладом квадратичной составляющей (далее LQ).

По результатам экспериментов при решении серий задач с существенной многоэкстремальностью (GRIS) было проведено корректное разделение на глобальные и локальные переменные, а затем задачи были успешно решены при помощи многошаговой схемы за приемлемое время и количество итераций. Задачи с функциями  GKLS для алгоритма являются более сложными, в силу схожести части параметров функций с параболоидами, поэтому решить все задачи серии за ограниченное количество итераций метода не удалось.

Вычисления проводились на компьютере с CPU Intel Core™ i7 10750H, 2.6 GHz. Для построения и анализа регрессионной модели использовался пакет scikit-learn из Python.
Численные результаты серий экспериментов приведены в Таблице \ref{tab1}.
В ней представлены данные о числе решенных задач $S$, среднем числе итераций глобального поиска на верхнем уровне рекурсии $G_{av}$, среднем числе испытаний на нижнем уровне $L_{av}$, среднем времени решения задачи $t_{av}$ (в секундах).

\begin{table}[ht]
	\caption{Результаты решения серий тестовых задач}
	\label{tab1}
	\begin{center}
		\begin{tabular}{ l c c c c c c } \hline
		 & $S$ &  $G_{av}$ &  $L_{av}$ & $t_{av}$ \\
    \hline
		GRIS-L & 20/20  & 705 &  378 545 & 1.3 \\
		GRIS-LQ & 20/20 & 705 &  420 397 & 1.4 \\
		GKLS-L & 18/20 & 9531 &  3 899 417 & 16.4 \\
		\hline
		\end{tabular}
	\end{center}
\end{table}

\begin{biblio}

\bibitem{Evtushenko2009} Евтушенко Ю. Г., Малкова В. У., Станевичюс А.-И. А. Параллельный поиск глобального экстремума функций многих переменных // {\it Журн. вычисл. матем. и матем. физ.} 2009. Т. 49. № 2. C. 255--269.

\bibitem{Paulavicius2016} Paulavi{\v c}ius R., {\v Z}ilinskas J. Advantages of Simplicial Partitioning for Lipschitz Optimization Problems with Linear Constraints // {\it Optim. Lett.} 2016. V. 10. No. 2. P. 237--246.

\bibitem{Grishagin2007} Городецкий С. Ю., Гришагин В. А. Нелинейное программирование и многоэкстремальная оптимизация. Н. Новгород: Изд-во ННГУ, 2007. 489 с.

\bibitem{Strongin13} Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А. Параллельные вычисления в задачах глобальной оптимизации. М.: Изд-во МГУ, 2013. 280 с.

\bibitem{Sergeyev2013} Sergeyev, Y.~D. and Strongin, R.~G. and Lera, D. Introduction to global optimization exploiting space-filling curves // {\it Journal of Global Optimization}, Springer. 2013. V. 60, No. 3. P. 595--596.

\bibitem{Grishagin2001} Sergeyev Y., Grishagin V. Parallel Asynchronous Global Search and the Nested Optimization Scheme //  {\it J. Comput. Anal. Appl.} 2001. V. 3, No. 2. P. 123--145.

%\bibitem{Sergeyev2008} Сергеев Я. Д., Квасов Д. Е. Диагональные методы глобальной оптимизации. М.: Физматлит, 2008. 352 с.

\end{biblio}

\newpage

\title{Separation of parameters using machine learning methods in global optimization problems}

\authors{K.A.~Barkalov,  M.A.~Usova}
\organizations{Lobachevsky State University of Nizhny Novgorod \\ Nizhny Novgorod, Russia}

% abstract is contained in the  abstract environment
\abstracten{
The paper presents the findings from research into an approach to solving global optimization problems with a different nature of dependence on diverse groups of parameters.
A scheme for choosing the problem parameters, which have a local effect on the objective function is proposed, which allows to solve essentially  multidimensional problems using the nested optimization scheme.
At the same time, different optimization algorithms are used at differing levels of recursion (nesting levels).
The proposed approach has demonstrated its efficiency in solving several series of test problems.
The effectiveness of the proposed approach was studied.}

\keywordsen{global optimization, local optimization, nested optimization scheme, high-dimensional problems, parameter separation}

\begin{biblioen}

\bibitem{Evtushenko2009} Evtushenko Yu. G., Malkova V. U., Stanevichyus A. A. Parallel global optimization of functions of several variables. // {\it USSR Computational Mathematics and Mathematical Physics} 2009. V. 49, No. 2. P. 255--269.

\bibitem{Paulavicius2016} Paulavi{\v c}ius R., {\v Z}ilinskas J. Advantages of Simplicial Partitioning for Lipschitz Optimization Problems with Linear Constraints // Optim. Lett. 2016. V. 10. No. 2. P. 237--246.

\bibitem{Grishagin2007} Gorodetskii S. Yu., Grishagin V. A. Nonlinear Programming and Optimization Multiextremal. Nizhny Novgorod, UNN Publ, 2007. 489 p.

\bibitem{Strongin13} Strongin, R.G., Gergel, V.P., Grishagin, V.A., Barkalov K.A. Parallel Computations in the Global Optimization Problems. Moscow, MSU Publ, 2012. 280 p.

\bibitem{Sergeyev2013} Sergeyev, Y.~D. and Strongin, R.~G. and Lera, D. Introduction to global optimization exploiting space-filling curves // Journal of Global Optimization, Springer. 2013. V. 60, No. 3. P. 595--596.

\bibitem{Grishagin2001} Sergeyev Y., Grishagin V. Parallel Asynchronous Global Search and the Nested Optimization Scheme // J. Comput. Anal. Appl. 2001. V. 3, No. 2. P. 123--145.

\end{biblioen}

\end{document}

### 2022-08-05 2022/Saransk/full_paper_Barkalov_Usova/paper.tex
add more information

@@
 
-Рассмотрим задачу оптимизации вида
+Задача многомерной многоэкстремальной оптимизации может быть определена как проблема поиска наименьшего значения действительной многоэкстремальной функции $\varphi(y)$
 \begin{eqnarray}\label{main_problem}
-& \varphi(y^\ast)=\min{\left\{\varphi(y): y\in D\right\}}, \nonumber \\
-& D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
+& \varphi(y^\ast)=\min{\left\{\varphi(y): y\in D\right\}},  \\
+& D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}, \nonumber
 \end{eqnarray}
-Будем предполагать, что функция $\varphi(y)$ является многоэкстремальной, заданной в виде черного ящика и удовлетворяет условию Липшица
+где $D$ есть область поиска, представляющая собой некоторый гиперпараллелепипед $N$-мерного евклидова пространства.
+Будем предполагать, что функция $\varphi(y)$ задана в виде черного ящика и удовлетворяет условию Липшица
 \[
@@
 
-Решение поставленной задачи можно организовать с помощью многошаговой схемы редукции размерности \cite{Grishagin2007}.
+Решение поставленной задачи можно организовать с помощью \textit{многошаговой схемы редукции размерности} \cite{Grishagin2007}, согласно которой решение многомерной задачи оптимизации может быть получено посредством решения последовательности «вложенных» одномерных задач.
+Таким образом, решение исходной задачи (\ref{main_problem}) можно проводить следующим образом:
+\begin{equation}\label{global_problem}
+\varphi(y^*) = \min_{y\in D} \varphi (y) = \min_{y^1\in D_1} f(y^1), \; D_1=\left\{y\in R^{N_1}: 
+a_i\leq y_i \leq b_i, \; 1\leq i \leq N_1\right\},
+\end{equation}
+где 
+\begin{equation}\label{local_problem}
+f(y^1) = \min_{ y^2 \in D_2} \varphi(y^1,y^2), \; D_2=\left\{y\in R^{N_2}: a_i\leq y_i \leq b_i, \; 
+N_1+1\leq i \leq N\right\}. 
+\end{equation}
+Данный подход широко используется для снижения сложности алгоритмов глобальной оптимизации и позволяет формировать эффективные покрытия области поиска.
+Решение  многоэкстремальной подзадачи (\ref{global_problem}), для которой требуется использовать сложные оптимизационные алгоритмы, будет проводиться на верхнем уровне рекурсии с помощью \textit{алгоритма глобального поиска} \cite{Strongin13}, принадлежащего классу характеристических алгоритмов, в сочетании со схемой редукции размерности на основе \textit{кривых Пеано} \cite{Sergeyev2013}.
+Этот алгоритм зарекомендовал себя как мощный инструмент для решения сложных многоэкстремальных задач.
 
-Решение  многоэкстремальной подзадачи, для которой требуется использовать сложные алгоритмы глобальной оптимизации, будет проводиться на верхнем уровне рекурсии с помощью алгоритма глобального поиска \cite{Strongin13} в сочетании со схемой редукции размерности на основе кривых Пеано \cite{Sergeyev2013}.
-Данный алгоритм зарекомендовал себя как мощный инструмент для решения сложных многоэкстремальных задач.
+Унимодальные подзадачи (\ref{local_problem}) на нижнем уровне могут быть решены с помощью методов прямого поиска, которые не предполагают гладкости функции.
+При поиске минимума эти методы измеряют только значения функций, потому в отличие от методов типа BFGS, они не требуют вычисления градиента и хорошо работают в условиях наличия вычислительных ошибок в значениях функций, что является типичным при решении прикладных задач.
+Одним из наиболее популярных в практике методов этого класса является \textit{метод Хука-Дживса}, использующийся в нашем алгоритме.
 
-Унимодальные подзадачи будут решаться на нижнем уровне с помощью метода Хука-Дживса.
-В отличие от методов типа BFGS, данный алгоритм не требует вычисления градиента и хорошо работает в условиях наличия вычислительных ошибок в значениях функций, что является типичным при решении прикладных задач.
-
-Данный подход поможет существенно снизить вычислительную сложность поиска оптимума.
+Описанный подход позволяет существенно снизить вычислительную сложность поиска оптимума.
+Однако его использование, вообще говоря, требует ручного распределения переменных по уровням, что, как уже отмечалось ранее, в практических задачах редко осуществимо.
+Потому возникла необходимость разработки схемы распределения параметров по уровням.
 
@@
 
-Если локальные переменные были обнаружены, то продолжить решение задачи с использованием описанной ранее многошаговой схемы.
-В противном случае запустить алгоритм глобального поиска, считая все переменные глобальными.
-
-Значения  $K$, $deg$ и $T$ являются параметрами метода и устанавливаются исследователем в соответствии со свойствами решаемой задачей.
-В реализованном методе $deg = 1$ или $deg = 2$, что позволяет построить модель линейной или квадратичной регрессии соответственно.
-
 \begin{figure}[h!]
 \centering \includegraphics[width=1\linewidth]{regression_94}
-\caption{Линейные регрессионные модели 15-мерной тестовой задачи  (2 глобальных параметра)}\label{regression_94_}
+\caption{Линейные регрессионные модели 20-мерной тестовой задачи  (2 глобальных параметра).}\label{regression_94_}
 \end{figure}
@@
 \centering \includegraphics[width=1\linewidth]{regression_94_3}
-\caption{Квадратичные регрессионные модели 15-мерной тестовой задачи (2 глобальных параметра)}\label{regression_94_3_}
+\caption{Квадратичные регрессионные модели 20-мерной тестовой задачи (2 глобальных параметра).}\label{regression_94_3_}
 \end{figure}
 
-В основе проводимой регрессии лежит метод наименьших квадратов (МНК).
-Оценка качества построенной на этапе 2 модели производится с помощью скорректированного коэффициента детерминации $R^2_a$.
-Указанный коэффициент характеризует схожесть построенной модели регрессии с целевой функцией в зафиксированном сечении.
+Значения  $K$, $deg$ и $T$ являются параметрами метода и устанавливаются исследователем в соответствии со свойствами решаемой задачей.
+В реализованном методе $deg = 1$ или $deg = 2$, что позволяет строить модель линейной или квадратичной регрессии соответственно.
+
+В основе проводимой регрессии лежит \textit{метод наименьших квадратов} (МНК).
+Оценка качества построенной модели производится с помощью скорректированного коэффициента детерминации $R^2_a$, характеризующего схожесть построенной модели регрессии с целевой функцией в зафиксированном сечении.
+
+\textbf{Этап 3.} Если локальные переменные были обнаружены, то продолжить решение задачи с использованием описанной ранее многошаговой схемы.
+В противном случае запустить алгоритм глобального поиска, считая все переменные глобальными.
 
@@
 
-Предложенный подход показал свою работоспособность при решении нескольких серий тестовых задач, представляющих собой линейные комбинации подзадач с глобальными параметрами (функции Гришагина или GRIS, функции Сергеева или GKLS \cite{Grishagin2001}) и подзадач с локальными параметрами, представляющих собой комбинацию близких к линейным одномерных функций (далее L) или  комбинацию  функций с разным вкладом квадратичной составляющей (далее LQ).
+Предложенный подход показал свою работоспособность при решении нескольких серий тестовых задач, представляющих собой линейные комбинации подзадач с глобальными параметрами и подзадач с локальными параметрами -- комбинации близких к линейным одномерных функций или функций с разным вкладом квадратичной составляющей.
 
-По результатам экспериментов при решении серий задач с существенной многоэкстремальностью (GRIS) было проведено корректное разделение на глобальные и локальные переменные, а затем задачи были успешно решены при помощи многошаговой схемы за приемлемое время и количество итераций.
+\subsection{Описание серий задач}
+
+В первом наборе тестовых задач (будем обозначать его GRIS-L) использовалось 20-мерные задачи комбинированного типа: локальная подзадача размерности 18, глобальная подзадача размерности 2.
+В качестве локальной составляющей использовались подзадачи, представляющие собой комбинацию близких к линейным одномерных функций:
+\begin{equation}\label{X2_problem}
+\varphi_j^{loc}(y) = \sum_{i=1}^{18} \left(M_{ij} y_i + L_{ij} y_i^2\right) ,
+\end{equation}
+где $-2.2 \leq y_i \leq 1.8$, а параметры $0.5 \leq |M_{ij}| \leq 1$ и $0.01 \leq L_{ij} \leq 0.3$  распределены случайно равномерно в соответствующих интервалах.
+В качестве глобальной составляющей задачи $\varphi_j^{glob}(y)$ использовались многоэкстремальные функции, предложенные Гришагиным \cite{Grishagin2001}.
+\begin{figure}[h!]
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1a.png}}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{fig1b.png}}
+\end{minipage}
+\caption{Линии уровня многоэкстремальной подзадачи и двумерное сечение локальной подзадачи.}
+\label{fig}
+\end{figure}
+\newpage
+Целевая функция $\varphi(y)$ во всех наборах представляла собой линейную комбинацию подзадач, т.е.
+$\varphi_j(y) = \alpha \varphi_j^{glob}(y) + \beta \varphi_j^{loc}(y)$.
+Последовательность параметров в каждой задаче задавалась случайным образом.
+
+Во втором наборе задач (далее GRIS-LQ) параметры $M_{ij}$ и $L_{ij}$ локальной подзадачи (\ref{X2_problem}) были устроены таким образом, чтобы в 60\% случаев переменная имела несущественную квадратичную часть, в 30\% вклад квадратичной составляющей был существенным (как минимум, соизмеримым со вкладом линейной составляющей), а оставшиеся 10\% переменных были чисто линейными.
+
+В третьей серии экспериментов (далее GKLS-L) в качестве многоэкстремальной части использовались 4-мерные функции из класса GKLS (функции Сергеева) \cite{Sergeyev2013} в качестве локальной составляющей использовалась локальные подзадачи серии GRIS-L.
+
+\subsection{Результаты экспериментов на тестовых сериях}
+
+По результатам экспериментов при решении серий задач с существенной многоэкстремальностью (GRIS-L и GRIS-LQ) было проведено корректное разделение на глобальные и локальные переменные, а затем задачи были успешно решены при помощи многошаговой схемы за приемлемое время и количество итераций.
+
 Задачи с функциями GKLS для алгоритма являются более сложными, в силу схожести части параметров функций с параболоидами, поэтому решить все задачи серии за ограниченное количество итераций метода не удалось.
@@
 Вычисления проводились на компьютере с CPU Intel Core™ i7 10750H, 2.6 GHz. Для построения и анализа регрессионной модели использовался пакет scikit-learn из Python.
+
 Численные результаты серий экспериментов приведены в Таблице \ref{tab1}.
@@
 \end{table}
+
+\subsection{Решение прикладной задачи}
+
+В РАЗРАБОТКЕ...
+
+\section{Заключение}
+
+В данной работе исследовался подход к разделению параметров по уровням для использования известной многоуровневой схемы редукции размерности в задачах глобальной оптимизации, комбинирующей использование кривых Пеано и многошаговую схему.
+Для решения редуцированных глобальных подзадач использовался алгоритм глобального поиска, для локальных подзадач был выбран метод Хука-Дживса.
+Предложена схема разделения параметров с использоваением построения регрессионной модели с помощью МНК.
+Проведены вычислительные эксперименты на серии тестовых задач большой размерности.
+Результаты экспериментов показывают, что предложенная схема позволяет корректно распределять параметры и эффективно находить оптимум задачи.
 

### 2022-08-05 2022/Saransk/full_paper_Barkalov_Usova/paper.tex
some fix in biblio
add some comments

@@
 При этом на разных уровнях рекурсии используются разные оптимизационные алгоритмы.
-Проведено исследование работоспособности предложенного подхода.
+Исследование работоспособности предложенного подхода проведено при решении тестовых и прикладных задач глобальной оптмизации.
 }
@@
 где $D$ есть область поиска, представляющая собой некоторый гиперпараллелепипед $N$-мерного евклидова пространства.
-Будем предполагать, что функция $\varphi(y)$ задана в виде черного ящика и удовлетворяет условию Липшица
+Будем предполагать, что функция $\varphi(y)$ задана в виде <<черного ящика>> и удовлетворяет условию Липшица
 \[
@@
 с априори неизвестной константой $L$.
+%Надо пояснить, что такое <<черный ящик>> - функция задается через алгоритм вычисления ее значений, аналитическая формула не известна.
+Алгоритм может быть сложным (решение СЛАУ или ОДУ), одно вычисление - трудоемкая операция.
+%Пояснить смысл условия Липшица для прикладных задач - ограниченное изменение аргумента порождает ограниченное изменение значений функции.
+
 Целевые функции большинства практических задач оптимизации обладают данным свойством.
@@
 \item сформировать множество $Q$, представляющее собой набор из $P+1$ пары вида $\left\{y_i^j, \varphi(z_i^j)\right\} $;
+%Здесь вместо deg лучше бы использовать какое-то буквенное обозначение. Вроде бы, M у нас нигде здесь не задействовано. То есть, можно говорить о полиномем степени M
 \item построить регрессионную модель заданной степени $deg$ на основе данных из $Q$;
@@
 \bibitem{Paulavicius2016} Paulavi{\v c}ius R., {\v Z}ilinskas J.
-Advantages of Simplicial Partitioning for Lipschitz Optimization Problems with Linear Constraints // Optim.
-Lett.
-2016.
+Advantages of Simplicial Partitioning for Lipschitz Optimization Problems with Linear Constraints // {\it Optim.
+Lett.} 2016.
 V.
@@
 and Lera, D.
-Introduction to global optimization exploiting space-filling curves // Journal of Global Optimization, Springer.
+Introduction to global optimization exploiting space-filling curves // {\it Journal of Global Optimization}, Springer.
 2013.
@@
 
-\bibitem{Grishagin2001} Sergeyev Y., Grishagin V. Parallel Asynchronous Global Search and the Nested Optimization Scheme // J. Comput. Anal. Appl. 2001. V. 3, No. 2. P. 123--145.
+\bibitem{Grishagin2001} Sergeyev Y., Grishagin V. Parallel Asynchronous Global Search and the Nested Optimization Scheme // {\it J. Comput. Anal. Appl.} 2001. V. 3, No. 2. P. 123--145.
 

### 2022-08-09 2022/PAVT Special Issue/template.tex
MPDI template

%  LaTeX support: latex@mdpi.com 
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[mathematics,article,submit,pdftex,moreauthors]{Definitions/mdpi} 
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, alloys, analytica, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, currophthalmol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, ent, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, futureinternet, futureparasites, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, kidneydial, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology, pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2022}
\copyrightyear{2022}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{} 
%\daterevised{} % Only for the journal Acoustics
\dateaccepted{} 
\datepublished{} 
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
\usepackage{inputenc}
\usepackage[T2A,T1]{fontenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}
%%%

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Using the ... Algorithm to Detect Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction}
% Дежурное название: Использование ...алгоритма для определения кинетических параметров процесса каталитической изомеризации пентан-гексановой фракции

% MDPI internal command: Title for citation in the left column
\TitleCitation{Title}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0001-5273-2471} % Add \orcidA{} behind the author's name
\newcommand{\orcidauthorB}{0000-0002-8736-0652} % Add \orcidB{} behind the author's name
\newcommand{\orcidauthorC}{0000-0003-4219-4870}

% Authors, for the paper (add full first names)
\Author{Konstantin Barkalov $^{1}$\orcidA{}, Irek Gubaydullin $^{2,3}$, Ilya Lebedev $^{1}$\orcidB{}, Roza Faskhutdinova $^{2,3}$ and Leniza Enikeeva $^{3,4,}$*\orcidC{}}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Konstantin Barkalov, Irek Gubaydullin, Ilya Lebedev, Roza Faskhutdinova and Leniza Enikeeva}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Barkalov, B.; Gubaydullin, I.; Lebedev, I.; Faskhutdinova, R.; Enikeeva, L.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

\address{%
$^{1}$ \quad Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia; konstantin.barkalov@itmm.unn.ru\\
$^{2}$ \quad Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia; irekmars@mail.ru \\
$^{3}$ \quad Ufa State Petroleum Technological University, Ufa, Russia; roza\_fask@mail.ru\\
$^{4}$ \quad Novosibirsk State University, Novosibirsk, Russia; leniza.enikeeva@yandex.ru
}

% Contact information of the corresponding author
\corres{Correspondence: leniza.enikeeva@yandex.ru (E.L.); konstantin.barkalov@itmm.unn.ru (B.K.) }

% Current address and/or shared authorship
%\firstnote{Current address: Affiliation 3.} 
%\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusions: indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{kinetic model; inverse problems of chemical kinetics; mathematical modeling; catalytic isomerisation of pantane-hexane fraction (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data
%\dataset{DOI number or link to the deposited data set if the data set is published separately. If the data set shall be published as a supplement to this paper, this field will be filled by the journal editors. In this case, please submit the data set as a supplement.}
%\datasetlicense{License under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia
%\encyclopediadef{For entry manuscripts only: please provide a brief overview of the entry title instead of an abstract.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

% Актуальность изучения процесса.
% В статье [L. V. Enikeeva, A. G. Faskhutdinov, K. F. Koledina, R. I. Faskhutdinova & I. M. Gubaydullin  Modeling and optimization of the catalytic isomerization of the pentane-hexane fraction with maximization of individual high-octane components yield. Reaction Kinetics, Mechanisms and Catalysis 133, 879–895 (2021).] мы рассматривали процесс, однако количество реакций необходимо сократить, потому что...
% Также в данной статье для расчета кинетических параметров применялся ген алгоритм и  время расчета кинетических параметров составляло около 10 часов. В данной статье мы будем использовать другой метод оптимизации, чтобы сократить времся расчета.

%Стыковочный абзац
Как правило, найти кинетические константы реакций аналитическим путем не представляется возможным, поэтому возникает необходимость в разработке и использовании численных методов их поиска (см., например, [ссылки (Наша совместная статья Суперкомп дни 2021 ??)]). В этом случае критерий качества найденного решения (целевая функция) не имеет явного аналитического описания, но допускает алгоритмическое представление и требует значительных затрат на расчет. Кроме того, в обратных задачах химической кинетики целевая функция может быть существенно многоэкстремальной, т.е. иметь много локальных экстремумов наряду с глобальным.

%Часть ННГУ
Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) являются вычислительно-трудоемкими и существенно отличаются от методов локального поиска (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}). Методы глобальной оптимизации можно условно разделить на два класса: метаэвристические и детерминированные. Метаэвристические алгоритмы, как правило, основаны на имитации процессов, протекающих в живой природе.
Примерами метаэвристических алгоритмов являются simulated annealing, evolution and genetic algorithms и т.д. (see, e.g., \cite{Battiti2009,Eiben2015}).
Указанные алгоритмы входят во многие библиотеки программных средств для решения задач глобальной оптимизации (например, MATLAB Global Optimization Toolbox, Scipy.Optimize и др. ).
В силу быстроты работы и малых требований к вычислительным ресурсам метаэвристические алгоритмы получили широкое распространение при решении практических задач.  Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может распологаться далеко от глобального \cite{Kvasov2018}.

Возможность построения детерминированных методов глобального поиска, отличающихся от brute force algorithm и метаэвристических алгоритмов, связана с наличием и учетом  априорных предположений о свойствах функций задачи. 
Одним из естественных допущений о задаче является предположение о том, что изменение значений целевой функции ограничено изменением ее аргумента. В таком случае, как известно, функции называют липшицевыми, а саму задачу --- задачей липшицевой глобальной оптимизации.

В данной статье отражены результаты применения параллельных методов липшицевой оптимизации, разрабатываемых в ННГУ им. Н.И. Лобачевского, к решению обратных задач химической кинетики. 
Основная часть статьи имеет следующую структуру. 
%Описание математической модели исследуемой химической реакции приведено в Section 2. Формальная постановка задачи липшицевой глобальной оптимизации приводятся в Section 3. В Section 4 кратко изложена схема разрабатываемого авторами асинхронного параллельного алгоритма для решения задач указанного класса. Результаты численного решения обратной задачи химической кинетики обсуждаются в Section 5.

\section{Mathematical model}

Let's consider the scheme of chemical transformations of the isomerization of the pentane-hexane fraction. The reaction consists of 48 reaction stages, ...

2-MP is 2-methyl pentane;

3-MP is 3-methyl pentane;

2,2-DMB is 2,2-dimethyl butane;

2,3-DMB is 2,3-dimethyl butane

CH* is cyclohexane

MCP is methyl cyclopentane

B is benzene.

3-MP is 3-methyl pentane;

2,2-DMB is 2,2-dimethyl butane;

2,3-DMB is 2,3-dimethyl butane

CH* is cyclohexane

\begin{table}[H] 
\caption{Transformations of the catalytic isomerization of the pentane-hexane fraction.\label{tab11}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCCC}
\toprule
\textbf{No}	& \textbf{Reaction}	& \textbf{E, kJ/mol}     & \textbf{log($k_0$)} \\
\midrule
1	& $n$-C$_5$H$_{12}$ $\rightarrow$ $iso$-C$_5$H$_{12}$ & 149.320 & 18.492\\
2 & $iso$-C$_5$H$_{12}$ $\rightarrow$ $n$-C$_5$H$_{12}$ & 160.474 & 19.507 \\
3 & $n$-C$_6$H$_{14}$ $\rightarrow$ 2-MP & 102.833 & 8.127 \\
4 & 2-MP $\rightarrow$ $n$-C$_6$H$_{14}$ & 106.317 & 11.279 \\

\bottomrule
\end{tabularx}
\end{table}

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}. Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.

%поправить название
\section{Parallel Algorithm for Solving Global Optimization Problems }\label{Sec_GSA}

\subsection{Постановка задачи глобальной оптимизации}

Как уже говорилось выше, с математической точки зрения the inverse problem of chemical kinetics может рассматриваться как Lipschitz global optimization problem. 
В общем виде задача указанного класса может быть сформулирована следующим образом:
\begin{gather}
 \phi^* = \phi(y^\ast)=\min{\left\{\phi(y):y\in D\right\}}, \label{problemN}\\
 D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\} \label{D},
\end{gather}
где $a,b \in R^N$ есть заданные границы области поиска, а целевая функция $\phi(y)$ удовлетворяет the Lipschitz condition
\begin{equation}\label{Lip}
\left|\phi(y_1)-\phi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D.
\end{equation}

Функция $\phi(y)$ предполагается многоэкстремальной и заданной в виде некоторого вычислительного алгоритма, на вход которого подается вектор параметров, а выходом является  соответствующее значение функции. Подобные функции часто называются функциями вида ``black box''. Также предполагается, что каждое вычисление значение функции в точке области поиска (называемое в дальнейшем также trial) является трудоемкой операцией и требует значительных вычислительных ресурсов. Как отмечалось во введении, такая постановка задачи полностью соответствует the inverse problem of chemical kinetics.

The Lipschitz condition (\ref{Lip}) может быть использовано для построения оценок глобального минимума функции в области поиска, на основе которых можно конструировать алгоритмы глобальной оптмизации и доказывать условия их сходимости (see, for example, \cite{Strongin2000}).

Одной из основных трудностей при решении Lipschitz global optimization problems является так называемое ``проклятие размерности'' --- экспоненциальный рост числа trials, требующегося для решения задачи с заданной точностью, при увеличении размерности задачи. Сгладить данный эффект при сохранении точности решения возможно за счет построения существенно неравномерных покрытий области поиска точками испытаний.

Известными примерами таких методов являются non-uniform space covering method \cite{Evtushenko2013}, simplicial partitions method \cite{Zilinskas2010}, diagonal partitions method {ссылка на Сергеева}. Указанные методы основаны на адаптивном разбиении области поиска на конечное число подобластей на каждой итерации.
Для указанных подобластей оценивается (с использованием накопленной информации о целевой функции) необходимость дальнейшего подразбиения для более детального исследования. 
 Принципиально иным подходом к решению многомерной задачи (\ref{problemN}) является ее редукция к одной или нескольким одномерным задачам с последующим применением эффективных методов одномерной оптмизации. Такая редукция может быть осуществлена, например, при помощи nested optimization scheme \cite{Grishagin2018} или при помощи Peano-Hilbert curves \cite{Barkalov2018}. Последний из указанных подходов и используется в настоящей работе.

Используя непрерывное однозначное отображение (Peano-Hilbert curve) $y(x)$ отрезка $[0,1]$ вещественной оси на гиперинтервал $D$ из (\ref{D}) можно свести многомерную задачу (\ref{problemN}) к одномерной задаче
\[
\phi(y^\ast)=\phi(y(x^\ast))=\min{\left\{\phi(y(x)): x\in[0,1]\right\}}.
\]
Известно \cite{Strongin2000,Sergeyev2013}, что редуцированная одномерная функция $\phi(y(x))$ уже не будет липшицевой, но будет удовлетворять H{\"o}lder condition
\[
\left|\phi(y(x_1))-\phi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}
\]
где H{\"o}lder constant $H$ связана с Lipschitz constant $L$ соотношением $ H=2 L \sqrt{N+3}$.

Конкретные способы численного построения подобного рода оботражений (называемых evolvents) рассмотрены в \cite{Strongin2000,Sergeyev2013}.
Здесь мы отметим, что evolvent будет аппроксимировать теоретическую Peano-Hilbert curve с точностью $2^{-m}$, где $m$ --- параметр. Примеры evolvent с разным значенем $m$ и $N$ приведены на Fig. 1.

Таким образом, в рамках данного подхода search trial в некотрой точке $x^k\in[0,1]$ будет включать в себя сначала the construction of the image $y^k=y(x^k)$, и лишь затем вычисление значения функции $ z^k = \phi(y^k)$.
Тройку значений $\{x^k,y^k,z^k\}$ мы будем называть outcome проведенного испытания.

%поправить название
\subsection{Асинхронный параллельный алгоритм поиска глобального минимума функции}

Используемый в данном исследовании параллельный алгоритм основан на информационно-статистический подходе к разработке методов глобальной оптимизации, изложенном в \cite{Strongin2000}. 
%сослаться на последовательный АГП

Нами была реализована асинхнонная схема распараллеливания вида ``master/worker''. В процессе-мастере выполняется алгоритм глобального поиска, в котором проводится накопление поисковой информации, оценка на ее основе константы Липшица для целевой функции, вычисление точек новых испытаний и распределение их по рабочим процессам. Процессы-рабочие получают от мастера точки, проводят в них новые испытания и отсылают мастеру результаты испытаний.

Будем считать, что на каждой итерации процесс-мастер вычисляет одну точку нового испытания и передает ее для проведения испытания в процесс-рабочий. При этом проведение испытания рабочим является значительно более трудоемкой операцией, чем выбор новой точки мастером, что исключает простой процессов-рабочих. 
В данном случае (в отличие от синхронных параллельных алгоритмов) общее число выполненных испытаний каждым процессом-рабочим будет зависеть от трудоемкости проведения конкретного испытания и не может быть оценено заранее.

При описании параллельного алгоритма будем предполагать, что есть один процесс-мастер и $p$ рабочих процессов. Верхним индексом будем обозначать порядковый номер проведенного испытания, а нижний индекс будем использовать для нумерации испытаний в порядке возрастания координаты.

Начальная итерация проводится по особому правилу. Процесс-мастер (будем считать, что это процесс No 0) инициирует параллельное проведение $p$ испытаний в $p$ различных точках области поиска, две из которых являются граничными, а остальные --- произвольными внутренними точками (например, распределенными равномерно). Таким оборазом, начальные испытания проводятся в точках $\{y(x^1), y(x^2), ...,y(x^p)\}$, где $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.

Предположим теперь, что процессом-мастером получены результаты $k$ испытаний, и в текущий момент времени процессы-рабочие проводят испытания в точках $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.

Пусть один из процессов (не ограничивая общности, можно считать, что это процессу No 1), в некоторый момент времени завершил проведение испытания в точке $y(x^{k+1})$.
Остальные процессы находятся на этапе проведения своих испытаний, т.е. в точках множества 
\[
I_k = \left\{ x^{k+2},...,x^{k+p} \right\}
\]
испытания уже начались, но еще не были завершены.

Тогда процесс No 1 пересылает процессу-мастеру полученные им результаты испытания, т.е. тройку  $\{x^{k+1},y^{k+1},z^{k+1}\}$.
Мастер сохраняет полученные результаты в своей информационной базе и выбирает для процесса No 1 точку нового испытания $x^{k+p+1}$ в соответствии с правилами, описанными ниже.

Step 1. Перенумеровать нижним идексом в порядке возрастания координаты точки множества
\[
X_k = \left\{x^1, x^2,...,x^{k+p} \right\},
\]
которое содержит все точки, в которых либо проведены, либо проводятся испытания, т.е.
\[
0=x_1<x_2<...<x_{x+p}=1.
\]
Step 2. Вычислить значения 
\[
M_1=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin I_k, x_i \notin I_k, 2\leq i\leq k+p \right\},
\]
\[
M_2=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in I_k, 2\leq i < k+p \right\},
\]
\[
M=\max\{M_1,M_2\},
\]
где $z_i=\phi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$. Значения $z_i$ в точках $x_i \in I_k$ являются неопределенными, т.к. испытания в точках $x_i \in I_k$ еще не завершены. Если значение $M$ получилось равным 0, то присвоить $M=1$.

Step 3. Для каждого интервала $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, вычислить значение 
\[
R(i)=rM\Delta_i+\frac{(z_i-z_{i-1})^2}{rM\Delta_i}-2(z_i+z_{i-1}),
\]
называемое характеристикой данного интервала. Здесь $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.

Step 4. Выбрать интервал $[x_{t-1},x_t]$ с наибольшим значением характеристики, т.е.
\[
R(t) = \max \left\{ R(i): \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p \right\}.
\]

Step 5. Вычислить точку $x^{k+p+1} \in (x_{t-1},x_t)$ в соответствии с формулой
\[
x^{k+p+1} = \frac{x_{t}+x_{t-1}}{2} - \mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_{t}-z_{t-1}\right|}{M}\right]^N.
\]

Сразу после вычисления точки $x^{k+p+1}$ очередного испытания процесс-мастер добавляет ее в множество $I_k$ и пересылает процессу-рабочему, который инициирует проведение испытания в ней.

Процесс-мастер завершает работу алгоритма, если выполняется либо условие $\Delta_{t}<\varepsilon$ ($t$ --- номер интервала c максимальной характеристикой из Правила 4), либо условие $k>K_{max}$ ($k$ --- число выполненных испытаний).
Вещественное число $\varepsilon>0$ и целое число $K_{max}>0$ являются параметрами алгоритма и соответствуют точности поиска решения и максимальному числу испытаний.

%Нужен комментарий про практическую реализацию алгоритма - используемые технологии, основные моменты реализации и т.п.

Отметим, что описанная здесь асинхронная схема распалаллеливания в отличие от синхронных схем, использовавшихся ранее при решении ряда прикладных задач (see \cite{Kalyulin2017,Modorskii2016}), обеспечивает полную загрузку всех задействованных процессов при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска. Это подтверждается результатами экспериментов, описанными в следующем разделе.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, Table~\ref{tab2}, etc.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
\caption{This is a figure. Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.\label{fig1}}
\end{figure}   
\unskip

\begin{table}[H] 
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCC}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data\\
\bottomrule
\end{tabularx}
\end{table}
\unskip

\begin{table}[H]
\caption{This is a wide table.\label{tab2}}
	\begin{adjustwidth}{-\extralength}{0cm}
		\newcolumntype{C}{>{\centering\arraybackslash}X}
		\begin{tabularx}{\fulllength}{CCCC}
			\toprule
			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
			\midrule
			Entry 1		& Data			& Data			& Data\\
			Entry 2		& Data			& Data			& Data \textsuperscript{1}\\
			\bottomrule
		\end{tabularx}
	\end{adjustwidth}
	\noindent{\footnotesize{\textsuperscript{1} This is a table footnote.}}
\end{table}

\subsection{Formatting of Mathematical Components}

This is the example 1 of equation:
\begin{linenomath}
\begin{equation}
a = 1,
\end{equation}
\end{linenomath}
the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

This is the example 2 of equation:
\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

% Example of a figure that spans the whole page width. The same concept works for tables, too.
\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=13.5cm]{Definitions/logo-mdpi}
\end{adjustwidth}
\caption{This is a wide figure.\label{fig2}}
\end{figure}

Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Patents}

%This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work~reported.}

\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''. Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}

\institutionalreview{Not applicable.}

\informedconsent{Not applicable.}

\dataavailability{In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study. Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}. If the study did not report any data, you might add ``Not applicable'' here.}

%\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections. This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}

\conflictsofinterest{The authors declare no conflict of interest.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\sampleavailability{Samples of the compounds ... are available from the authors.}

%% Only for journal Encyclopedia
%\entrylink{The Link to this entry published on the encyclopedia platform.}

%\abbreviations{Abbreviations}{
%The following abbreviations are used in this manuscript:\\
%
%\noindent 
%\begin{tabular}{@{}ll}
%MDPI & Multidisciplinary Digital Publishing Institute\\
%DOAJ & Directory of open access journals\\
%TLA & Three letter acronym\\
%LD & Linear dichroism
%\end{tabular}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{adjustwidth}{-\extralength}{0cm}
%\printendnotes[custom] % Un-comment to print a list of endnotes

\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
\bibliography{bibliography}

%=====================================
% References, variant B: internal bibliography
%=====================================

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{adjustwidth}
\end{document}

### 2022-08-11 2022/Saransk/full_paper_Barkalov_Usova/paper.tex
add results by UFA problem

@@
 \end{figure}
-\newpage
+%\newpage
 Целевая функция $\varphi(y)$ во всех наборах представляла собой линейную комбинацию подзадач, т.е.
@@
 
-В РАЗРАБОТКЕ...
+Текущее изменение климата, вызывающее опасные геологические природные явления всё чаще, определяет множество актуальных прикладных задач.
+К таким задачам относится изучение схода оползней, селей и лавин, их предсказания и обнаружения.
+В России подобные опасные явления характерны для Краснодарского края и Северного Кавказа.
+Моделирование схода склонового потока позволяет предсказать разрушения вызванные этим явлением, грамотно расположить защитные сооружения и жизненно-важные объекты.
+
+Подробности описания турбулентной модели для склоновых потоков представлены в работах \cite{} ?.
+При помощи глобальной оптимизации подбираются коэффициенты модели (размерность задачи 15).
+Ранее с помощью параллельного асинхронного алгоритма, работающего на 8 узлах кластера, были получены оптимальные для склоновых потоков коэфициенты турбулентности при параметрах метода  $r = 3$ и $\epsilon = 10^{-3}$.
+За [сколько????] часов было найдено лучшее значение целевой функции $0.519793$.
+Оптимизируемые (главные) параметры определялись из физического смысла задачи.
+
+С помощью исследуемого в данной работе метода за 15 часов было найдено улучшенное решение данной задачи -- $0.497063$.
+Размерность глобальной подзадачи -- $4$, число запусков глобального поиска -- $427$, число запусков локального метода -- [добавила в optim.dat запись локальных запусков, можно будет позже записать].
+
+\begin{figure}[h!]
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{saved_figure_2.png}}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{saved_figure_8.png}}
+\end{minipage}
+\caption{Квадратичные регрессионные модели 15-мерной прикладной задачи глобального параметра (слева) и для локального параметра (справа).}
+\label{fig}
+\end{figure}
 
@@
 Проведены вычислительные эксперименты на серии тестовых задач большой размерности.
+Получено улучшенное решение прикладной задачи.
 Результаты экспериментов показывают, что предложенная схема позволяет корректно распределять параметры и эффективно находить оптимум задачи.

### 2022-08-12 2022/PAVT Special Issue/template.tex
new figure

@@
 %Стыковочный абзац
-Как правило, найти кинетические константы реакций аналитическим путем не представляется возможным, поэтому возникает необходимость в разработке и использовании численных методов их поиска (см., например, [ссылки (Наша совместная статья Суперкомп дни 2021 ??)]).
-В этом случае критерий качества найденного решения (целевая функция) не имеет явного аналитического описания, но допускает алгоритмическое представление и требует значительных затрат на расчет.
+Как правило, найти кинетические константы реакций аналитически не представляется возможным, поэтому возникает необходимость в разработке и использовании численных методов их поиска (см., например, [ссылки]).
+В этом случае критерий качества найденного решения (objective function) задается в виде некоторого алгоритма, который включает в себя этап численного моделирования и, как следствие, является трудоемким.
 Кроме того, в обратных задачах химической кинетики целевая функция может быть существенно многоэкстремальной, т.е.
@@
 %Часть ННГУ
-Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) являются вычислительно-трудоемкими и существенно отличаются от методов локального поиска (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}).
-Методы глобальной оптимизации можно условно разделить на два класса: метаэвристические и детерминированные.
+Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) существенно отличаются от методов локальной оптимизации, т.к.
+требуют исследования всей search domain (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}).
+Указанные методы можно условно разделить на два класса: метаэвристические и детерминированные.
 Метаэвристические алгоритмы, как правило, основаны на имитации процессов, протекающих в живой природе.
@@
 Указанные алгоритмы входят во многие библиотеки программных средств для решения задач глобальной оптимизации (например, MATLAB Global Optimization Toolbox, Scipy.Optimize и др. ).
-В силу быстроты работы и малых требований к вычислительным ресурсам метаэвристические алгоритмы получили широкое распространение при решении практических задач.
+В силу быстроты работы и малых требований к вычислительным ресурсам (в частности, памяти) метаэвристические алгоритмы получили широкое распространение при решении практических задач.
 Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может распологаться далеко от глобального \cite{Kvasov2018}.
@@
 Одним из естественных допущений о задаче является предположение о том, что изменение значений целевой функции ограничено изменением ее аргумента.
-В таком случае, как известно, функции называют липшицевыми, а саму задачу --- задачей липшицевой глобальной оптимизации.
+В таком случае, как известно, функции называют липшицевыми, а саму задачу --- Lipschitz global optimization problem.
 
@@
 Конкретные способы численного построения подобного рода оботражений (называемых evolvents) рассмотрены в \cite{Strongin2000,Sergeyev2013}.
-Здесь мы отметим, что evolvent будет аппроксимировать теоретическую Peano-Hilbert curve с точностью $2^{-m}$, где $m$ --- параметр. Примеры evolvent с разным значенем $m$ и $N$ приведены на Fig. 1.
+Здесь мы отметим, что evolvent будет аппроксимировать теоретическую Peano-Hilbert curve с точностью $2^{-m}$, где $m$ --- параметр, задающий точность аппроксимации.
+Примеры evolvent для разных размерностей приведены на Fig.~\ref{fig:Peano}.
+
+\begin{figure}
+\center
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=0.9\linewidth]{fig1a.JPG} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=0.9\linewidth]{fig1b.JPG} \\ (b)}
+\end{minipage}
+\caption{Evolvents with $m=4$ and (a) $N=2$, (b) $N=3$.}\label{fig:Peano}
+\end{figure}
 
@@
 
-Используемый в данном исследовании параллельный алгоритм основан на информационно-статистический подходе к разработке методов глобальной оптимизации, изложенном в \cite{Strongin2000}. 
-%сослаться на последовательный АГП
+Используемый в данном исследовании параллельный алгоритм основан на информационно-статистическом подходе к разработке методов глобальной оптимизации, теоретические основы которого изложены в \cite{Strongin2000}.
 
@@
 
-Пусть один из процессов (не ограничивая общности, можно считать, что это процессу No 1), в некоторый момент времени завершил проведение испытания в точке $y(x^{k+1})$.
-Остальные процессы находятся на этапе проведения своих испытаний, т.е. в точках множества 
+Пусть один из процессов (не ограничивая общности, можно считать, что это процесс No 1), в некоторый момент времени завершил проведение испытания в точке $y(x^{k+1})$.
+Остальные процессы находятся в стадии проведения своих испытаний, т.е. в точках множества 
 \[
@@
 
+Последовательный алгоритм глобального поиска, на основе которого была реализованна предложенный асинхронный алгоритм, допускает следующую интерпретацию \cite{Molinaro2001}.
+Используя точки ранее проведенных испытаний можно построить миноранту целевой функции, для которой значение $-R(i)$, где $R(i)$ --- значение характеристики интервала $(x_{i-1},x_i)$, будет являться значением миноранты в точке ее миинмума на интервале $(x_{i-1},x_i), 1\leq i\leq k+p$.
+В соответствии с правилом 4 алгоритма, новое испытание проводится в интервале с наименьшим значением миноранты.
+А в соответствии с правилом 5 точка нового испытания совпадает с точкой минимума миноранты.
+
 %Нужен комментарий про практическую реализацию алгоритма - используемые технологии, основные моменты реализации и т.п.
+%Асинхронный алгоритм глобального поиска был реализован с помощью технологии MPI, язык разработки --- С++.
 

### 2022-08-12 2022/Saransk/full_paper_Barkalov_Usova/paper.tex
add results by chimic problem

--- 2022-08-05 2022/Saransk/full_paper_Barkalov_Usova/paper.tex
@@
 
-Вычисления проводились на компьютере с CPU Intel Core™ i7 10750H, 2.6 GHz. Для построения и анализа регрессионной модели использовался пакет scikit-learn из Python.
+Вычисления проводились на компьютере с CPU Intel Core™ i7 10750H, 2.6 GHz.
+Исследуемый алгоритм был реализован на C++; вычисление значений целевой функции было реализовано в Python 3.9.
+Для построения и анализа регрессионной модели использовался пакет scikit-learn из Python.
 
@@
 
-В РАЗРАБОТКЕ...
+В настоящее время наблюдается тенденция к улучшению экологических характеристик автомобильного топлива при сохранении высокого октанового числа.
+Алкилирование изобутана смешанными олефинами, катализируемое серной кислотой, позволяет получить высокооктановый компонент бензина с минимальным содержанием ароматических углеводородов.
+Для оптимизации описанного химического процесса в промышленности необходимо сначала разработать его модель, что в данном случае означает построение математической модели химического процесса, а затем его кинетической модели, т.
+е.
+численно рассчитать кинетические константы реакции.
+
+Подробные описания моделей представлены в работе \cite{KinModel2022}.
+
+Аналитически узнать кинетические константы реакций, как правило, невозможно.
+В этом случае критерии качества найденного решения (целевая функция) не имеют явного аналитического описания, но допускают алгоритмическое представление и требуют значительных вычислительных ресурсов.
+Ранее использование параллельного асинхронного алгоритма, работающего на 8 узлах кластера, позволило значительно сократить время поиска оптимума.
+Для проведения указанного численного эксперимента использовался UNN supercomputer Lobachevsky (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node).
+Найденные оптимальные параметры 15-мерной модели позволили адекватно смоделировать описанный процесс.
+При параметре метода  $r = 3$ было найдено лучшее значение целевой функции $0.519793$.
+Однако при использовании данного метода решения достигнуть заданной точности за обозримое время не представляется возможным (используется остановка по числу итераций или по ограничению времени решения).
+
+С помощью исследуемого в данной работе метода за 15 часов было найдено улучшенное решение данной задачи -- $0.497063$.
+Размерность глобальной подзадачи -- $4$, число запусков глобального поиска -- $427$, число запусков локального метода --$427 000$.
+
+\begin{figure}[h!]
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{saved_figure_2.png}}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+\center{\includegraphics[width=1.0\linewidth]{saved_figure_8.png}}
+\end{minipage}
+\caption{Квадратичные регрессионные модели 15-мерной прикладной задачи глобального параметра (слева) и для локального параметра (справа).}
+\label{fig}
+\end{figure}
 
@@
 \bibitem{Grishagin2001} Sergeyev Y., Grishagin V. Parallel Asynchronous Global Search and the Nested Optimization Scheme //  {\it J. Comput. Anal. Appl.} 2001. V. 3, No. 2. P. 123--145.
+
+\bibitem{KinModel2022} Gubaydullin, I., Enikeeva, L., Barkalov, K., Lebedev, I., Silenko, D.
+Kinetic Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as a Catalyst Using the Asynchronous Global Optimization Algorithm.
+// {\it Communications in Computer and Information Science}.
+2022.
+V.
+1618.
+P.
+293--306.
 
@@
 
-\bibitem{Sergeyev2013} Sergeyev, Y.~D.
-and Strongin, R.~G.
+\bibitem{Sergeyev2013} Sergeyev, Y.D.
+and Strongin, R.G.
 and Lera, D.
@@
 
+\bibitem{KinModel2022} Gubaydullin, I., Enikeeva, L., Barkalov, K., Lebedev, I., Silenko, D.
+Kinetic Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as a Catalyst Using the Asynchronous Global Optimization Algorithm.
+// {\it Communications in Computer and Information Science}.
+2022.
+V.
+1618.
+P.
+293--306.
+
 \end{biblioen}

### 2022-08-12 2022/Saransk/full_paper_Barkalov_Usova/paper.tex
Update paper.tex
fix some misprints

@@
 При этом на разных уровнях рекурсии используются разные оптимизационные алгоритмы.
-Исследование работоспособности предложенного подхода проведено при решении тестовых и прикладных задач глобальной оптмизации.
+Исследование работоспособности предложенного подхода проведено при решении тестовых и прикладных задач глобальной оптимизации.
 }
@@
 
-\subsection{Результаты экспериментов на тестовых сериях}
+\subsection{Результаты экспериментов на сериях тестовых задач}
 
@@
 Алкилирование изобутана смешанными олефинами, катализируемое серной кислотой, позволяет получить высокооктановый компонент бензина с минимальным содержанием ароматических углеводородов.
-Для оптимизации описанного химического процесса в промышленности необходимо сначала разработать его модель, что в данном случае означает построение математической модели химического процесса, а затем его кинетической модели, т.
-е.
+Для оптимизации описанного химического процесса в промышленности необходимо сначала разработать его модель, что в данном случае означает построение математической модели химического процесса, а затем его кинетической модели, т.е.
 численно рассчитать кинетические константы реакции.
@@
 Ранее использование параллельного асинхронного алгоритма, работающего на 8 узлах кластера, позволило значительно сократить время поиска оптимума.
-Для проведения указанного численного эксперимента использовался UNN supercomputer Lobachevsky (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node).
+Для проведения указанного численного эксперимента использовался суперкомпьютер <<Лобачевский>> (операционная система -- CentOS 7.2, на узле -- два процессора Intel Sandy Bridge E5-2660 2.2 GHz, оперативная память -- 64 Gb).
 Найденные оптимальные параметры 15-мерной модели позволили адекватно смоделировать описанный процесс.
@@
 С помощью исследуемого в данной работе метода за 15 часов было найдено улучшенное решение данной задачи -- $0.497063$.
-Размерность глобальной подзадачи -- $4$, число запусков глобального поиска -- $427$, число запусков локального метода --$427 000$.
+Размерность глобальной подзадачи -- $4$, число запусков глобального поиска -- $427$, число запусков локального метода -- $427 000$.
 
@@
 Для решения редуцированных глобальных подзадач использовался алгоритм глобального поиска, для локальных подзадач был выбран метод Хука-Дживса.
-Предложена схема разделения параметров с использоваением построения регрессионной модели с помощью МНК.
+Предложена схема разделения параметров с использованием построения регрессионной модели с помощью МНК.
 Проведены вычислительные эксперименты на серии тестовых задач большой размерности.

### 2022-08-16 2022/PAVT Special Issue/template.tex
correct UNN introduction

@@
 
-% Актуальность изучения процесса.
-% В статье [L.
-V.
-Enikeeva, A.
-G.
-Faskhutdinov, K.
-F.
-Koledina, R.
-I.
-Faskhutdinova & I.
-M.
-Gubaydullin  Modeling and optimization of the catalytic isomerization of the pentane-hexane fraction with maximization of individual high-octane components yield.
-Reaction Kinetics, Mechanisms and Catalysis 133, 879–895 (2021).] мы рассматривали процесс, однако количество реакций необходимо сократить, потому что...
-% Также в данной статье для расчета кинетических параметров применялся ген алгоритм и  время расчета кинетических параметров составляло около 10 часов.
-В данной статье мы будем использовать другой метод оптимизации, чтобы сократить времся расчета.
-
-%Стыковочный абзац
-Как правило, найти кинетические константы реакций аналитически не представляется возможным, поэтому возникает необходимость в разработке и использовании численных методов их поиска (см., например, [ссылки]).
-В этом случае критерий качества найденного решения (objective function) задается в виде некоторого алгоритма, который включает в себя этап численного моделирования и, как следствие, является трудоемким.
+%Стыковочный абзац - его надо переписать
+В сложных задачах химической кинетики найти кинетические константы реакций аналитически не представляется возможным, поэтому возникает необходимость в применеии численных методов для их поиска (см., например, [ссылки]).
+В этом случае критерий качества (objective function) задается в виде некоторого алгоритма, который включает в себя этап численного моделирования и, как следствие, является трудоемким.
 Кроме того, в обратных задачах химической кинетики целевая функция может быть существенно многоэкстремальной, т.е.
@@
 
-%Часть ННГУ
+%Часть ННГУ - начало
 Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) существенно отличаются от методов локальной оптимизации, т.к.
-требуют исследования всей search domain (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}).
+требуют исследования всей search domain (see, e.g., \cite{PaulaviciusZilinskas2014,Sergeyev2017}).
 Указанные методы можно условно разделить на два класса: метаэвристические и детерминированные.
@@
 В силу быстроты работы и малых требований к вычислительным ресурсам (в частности, памяти) метаэвристические алгоритмы получили широкое распространение при решении практических задач.
-Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может распологаться далеко от глобального \cite{Kvasov2018}.
+Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может распологаться далеко от глобального оптимума \cite{Kvasov2018}.
 
@@
 В таком случае, как известно, функции называют липшицевыми, а саму задачу --- Lipschitz global optimization problem.
+Центральным направлением развития современных методов липшицевой оптимизации является разработка параллельных алгоритмов.
+При этом предполагается, что значения целевой функции вычисляются в результате решения какой-либо другой вычислительной задачи и требуют проведения затратного вычислительного эксперимента.
 
-В данной статье отражены результаты применения параллельных методов липшицевой оптимизации, разрабатываемых в ННГУ им. Н.И. Лобачевского, к решению обратных задач химической кинетики. 
+В данной статье отражены результаты применения параллельных методов липшицевой оптимизации, развиваемых в работах [], к решению inverse problems of chemical kinetics. 
 Основная часть статьи имеет следующую структуру. 
+%Часть ННГУ - конец
+
 %Описание математической модели исследуемой химической реакции приведено в Section 2.
@@
 
+%Часть ННГУ - начало
 %поправить название
@@
 
-Известными примерами таких методов являются non-uniform space covering method \cite{Evtushenko2013}, simplicial partitions method \cite{Zilinskas2010}, diagonal partitions method {ссылка на Сергеева}.
+Известными примерами таких методов являются non-uniform space covering method \cite{Evtushenko2013}, simplicial partitions method \cite{PaulaviciusZilinskas2014}, diagonal partitions method \cite{Sergeyev2017}.
 Указанные методы основаны на адаптивном разбиении области поиска на конечное число подобластей на каждой итерации.
@@
 
+Теория сходимости и различные модификации последовательных и параллельных алгоритмов детально изложены в монографиях \cite{Strongin2000,Sergeyev2017}. 
+Отметим, что описанная здесь асинхронная схема распалаллеливания в отличие от синхронных схем, использовавшихся ранее при решении ряда прикладных задач (see \cite{Kalyulin2017,Modorskii2016}), обеспечивает полную загрузку всех задействованных процессов при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска.
+
 %Нужен комментарий про практическую реализацию алгоритма - используемые технологии, основные моменты реализации и т.п.
+
 %Асинхронный алгоритм глобального поиска был реализован с помощью технологии MPI, язык разработки --- С++.
 
-Отметим, что описанная здесь асинхронная схема распалаллеливания в отличие от синхронных схем, использовавшихся ранее при решении ряда прикладных задач (see \cite{Kalyulin2017,Modorskii2016}), обеспечивает полную загрузку всех задействованных процессов при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска.
-Это подтверждается результатами экспериментов, описанными в следующем разделе.
+%Часть ННГУ - конец
 
@@
 \end{figure}
-
-Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
-%% Example of a theorem:
-\begin{Theorem}
-Example text of a theorem.
-\end{Theorem}
-
-The text continues here. Proofs must be formatted as follows:
-
-%% Example of a proof:
-\begin{proof}[Proof of Theorem 1]
-Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
-\end{proof}
-The text continues here.
 

### 2022-08-18 2022/OPTIMA/paper.tex
fix the most part of the comments

@@
 \begin{equation}\label{Q}
-Q = \left\{x \in R^N : a_j \leq x_j \leq b_j, 1 \leq i \leq N \right\}
+Q = \left\{x \in R^N : a_j \leq x_j \leq b_j, 1 \leq j \leq N \right\}
 \end{equation}
@@
 
-For solving the problems of the family, a global optimization technique based on reducing the multidimensional problem (\ref{conv_problem}) to an equivalent univariate subproblem \cite{Strongin2000,Pinter1996,Zhigljavsky2008,Sergeyev2013,PaulaviciusZilinskas2014,Sergeyev2017} in combination with one-dimensional information algorithm with accelerated convergence is proposed.
+For solving the problems of the family, a global optimization technique based on reducing the multidimensional problem (\ref{conv_problem}) to an equivalent univariate subproblem \cite{Strongin2000,Pinter1996,Zhigljavsky2008,Sergeyev2013,PaulaviciusZilinskas2014,Sergeyev2017}
+in combination with one-dimensional information algorithm with accelerated convergence is proposed.
 
@@
 
-It is known as a fundamental fact that $N$-dimensional hyperparallelepiped (\ref{Q}) and the interval $[0,1]$ of the real axis are the equipotent sets and the interval $[0,1]$ can be mapped onto the parallelepiped (\ref{Q}) unambiguously and continuously.
+It is known as a fundamental fact (see, e.g., \cite{Sagan1994}) that $N$-dimensional hyperparallelepiped (\ref{Q}) and the interval $[0,1]$ of the real axis are the equipotent sets and the interval $[0,1]$ can be mapped onto the parallelepiped (\ref{Q}) unambiguously and continuously.
 Such mappings are called \textit{Peano-type curves} or \textit{evolvents}.
 
-Let $x(t), \; t \in [0,1]$  be a Peano-type curve, and the function  $f(x)$ from (\ref{problem_f}) be continuous. As
+Let $x(t), \; t \in [0,1],$  be a Peano-type curve, and the function  $f(x)$ from (\ref{problem_f}) be continuous. As
 \[
@@
 
+%However, the function $f(x(t))$ will not meet the Lipschitz if $f(x)$ is Lipschitzian. It will satisfy the H{\"o}lder condition
 However, in the case when the function $f(x)$ is Lipschitzian this property is not valid for the reduced function. The function $f(x(t))$ will satisfy the H{\"o}lder condition
@@
 \end{equation}
-with a H{\"o}lder constant $H>0$ \cite{Strongin2000,Sergeyev2013}.
-
-If the function $f(x)$ satisfies the Lipschitz condition (\ref{lip_1}) then $H=2L\sqrt{N+3}$.
+with a H{\"o}lder constant $H>0$.
+And if the function $f(x)$ satisfies the Lipschitz condition (\ref{lip_1}) then $H=2L\sqrt{N+3}$ (see \cite{Sergeyev2013,Strongin2000}).
 As a consequence, minimization of the function $f(x(t))$  requires the use of special algorithms oriented at functions with property (\ref{holder}).
@@
 
-\begin{enumerate}
-\item 
+Step 1.
 The points $t^1, \dots, t^n$ of completed trials are ordered in increasing order and renumbered by subscripts, i.e.,
@@
 The values $v_s = \varphi(t_s)$ are juxtaposed to the points  $t_s$, $1 \leq s \leq n$.
-\item 
+
+Step 2. 
 A real values $R(q)$ are assigned to all the intervals $(t_{q-1}, t_q)$, $2 \leq q \leq n$, formed by neighboring points from (\ref{eq:15}).
@@
 Here
-\begin{equation}\label{eq:18}
-\begin{matrix}
+\begin{gather}\label{eq:18}
 B(q) = (t_s - t_{s-1})^{1/N} + \frac{(v_q - v_{q-1})^2} {m^2 (t_s - t_{s-1})^{1/N}} -2\frac{v_{q} + v_{q-1}}{m},\\
-\varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
-\end{matrix}
-\end{equation}
+\varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)} \nonumber
+\end{gather}
 is minimal computed value of the objective function $\varphi(t)$ at trial points,
@@
 
-\item
+Step 3. 
 Among all intervals, the interval   with the highest characteristic is selected
@@
 
-\item
+Step 4.
 The new $(n+1)$-th trial is performed at the point
@@
 
-\end{enumerate}
-
-The method ALR described above is a modification of the basic information-statistical algorithm \cite{Strongin2000,Sergeyev2013} for minimization functions satisfying the H{\"o}lder condition.
+The method ALR described above is a modification of the basic information-statistical algorithm \cite{Sergeyev2013,Strongin2000} for minimization functions satisfying the H{\"o}lder condition.
 The modification consists in introducing new characteristics (\ref{eq:17}) after initial stage of the search to improve the method's speedup on account of increasing values of characteristics for intervals containing current record (\ref{eq:18}).
@@
 
-Like its prototype, ARL can apply as the stopping condition the inequality
+Like its prototype \cite{Sergeyev2013,Strongin2000}, ARL can apply as the stopping condition the inequality
 \begin{equation}\label{eq:23}
@@
 
-The complete convergence theory of the information-statistical algorithms is given in \cite{Strongin2000,Sergeyev2013}.
+The complete convergence theory of the information-statistical algorithms (including proof of the statements (\ref{eq:24}) and (\ref{eq:25})) is given in \cite{Sergeyev2013,Strongin2000}.
 
@@
 
-In the course of optimization the algorithm builds the trial points $t^1,...,t^n$ which correspond to the points $x^1 = x(t^1),...,x^n=x(t^n)$ in the domain $Q$ .
+In the course of optimization the algorithm generates the trial points $t^1,...,t^n$ which correspond to the points $x^1 = x(t^1),...,x^n=x(t^n)$ in the domain $Q$ .
 At these points the values $v^j = \Phi_\lambda(x^j), \; 1 \leq j \leq n,$ of the function $\Phi_\lambda(x)$ from (\ref{conv_problem}) are computed as values $\varphi(t^j)$ of the function $\varphi(t)$  minimized by ALR.
@@
 Since all GKLS functions have a global minimum value equal to $-1$, the partial  criteria $w_i(x), \; 1 \leq i \leq k,$  were increased by a positive constant to ensure the requirements (\ref{positive}).
-
-In all experiments the domain
-\[
-Q = \left\{x \in R^N : -1 \leq x_j \leq -1, 1 \leq i \leq N \right\}
-\]
-was considered as the feasible one.
+In all experiments the domain $Q$ from (\ref{Q}) with $a_j=-1, \; b_j=1$ was considered as the feasible one.
 

### 2022-08-18 2022/PAVT Special Issue/template.tex
Добавлен текст про реализацию

@@
 
+\subsection{Особенности программной реализации параллельного алгоритма}
+
+\begin{figure}
+\center{\includegraphics[width=0.9\linewidth]{fig2.JPG}}
+\caption{Схема параллельных вычислений в системе Globalizer \\ (1 -- MPI связь, 2 -- связь через потоки).}\label{fig:Impl}
+\end{figure}
+
+В нашей реализации алгоритма глобального поиска для организации параллельных вычислений используется MPI.
+Процесс с рангом 0 выполняет роль мастера.
+В рамках процесса реализован асинхронный алгоритм глобального поиска (см.
+раздел 3.2 шаги 1-5).
+Процессы с рангом отличным от 0 выполняют роль вычислителей (рабочих).
+
+Принцип взаимодействия процессов (рисунок \ref{fig:Impl}, связь 1) заключается в следующем.
+В рамках первого шага на мастере выполняется выбор $p$ точек для проведения испытаний (см.
+раздел 3.2).
+Выбранные точки проведения испытаний рассылаются вычислителям и мастер переходит в состояние ожидания приема сообщений.
+Сообщения от вычислителей содержат outcome проведенного испытания.
+Полученная информация используется для выбора точки очередного испытания.
+Новая точка отправляется в виде сообщения на вычислитель.
+Если в GSA выполнено условие остановки, то вместо точки для испытания на вычислитель отправляется сообщение о завершении работы.
+
+Каждый процесс, выполняющий роль вычислителя, при старте создает экземпляр задачи GSA, а затем переходит в состояние ожидания сообщений от мастер-процесса.
+При приеме сообщения содержащего точку испытания, выполняется вычислительный эксперимент.
+Результаты эксперимента отправляются мастеру.
+После проведения эксперимента сам процесс продолжает ожидание новых сообщений.
+При приеме сообщения о завершении вычислений выполняется высвобождение ресурсов, используемых задачей GSA, и завершение основного процесса.
+
+При реализации асинхронной схемы мы предполагали, что время начала вычисления значений функции критерия на разных процессах отличается.
+Наша реализация учитывает тот факт, что точка испытания, а также результаты вычислений представляют незначительный объем данных, который относительно быстро можно отправить по сети.
+Как результат, для взаимодействия мастера и вычислителей (рисунок \ref{fig:Impl}, связи 1) использовались базовые операции пересылки сообщений MPI типа точка-точка.
+
+При проведении вычислительных экспериментов (см.
+раздел 5) задача химической кинетики была реализована в рамках MATLAB.
+В исходном варианте задачи каждый набор параметров требовал запуск процесса MATLAB, передачу точки проведения испытаний, а также получение результатов вычислительных экспериментов.
+Для сокращения накладных расходов на организацию вычислений были выполнены ряд оптимизаций.
+Во-первых, исходный код для вычислительного эксперимента был скомпилирован в исполняемое приложение.
+Во-вторых, был реализован подход, при котором MATLAB код выполнял ожидание очередной точки испытаний через входной поток, а через выходной поток осуществлял возврат результатов вычислительных экспериментов.
+Подобный подход позволил в каждом вычислителе один раз запустить процесс с откомпилированным MATLAB кодом, перехватить входные и выходные потоки, а затем выполнять вычисления без ожидания запуска новых процессов.
+
 %Часть ННГУ - конец

### 2022-08-18 2022/PAVT Special Issue/template.tex
correct subsection 3.3

@@
 
-Последовательный алгоритм глобального поиска, на основе которого была реализованна предложенный асинхронный алгоритм, допускает следующую интерпретацию \cite{Molinaro2001}.
+Последовательный алгоритм глобального поиска, на основе которого был реализован предложенный асинхронный алгоритм, допускает следующую интерпретацию \cite{Molinaro2001}.
 Используя точки ранее проведенных испытаний можно построить миноранту целевой функции, для которой значение $-R(i)$, где $R(i)$ --- значение характеристики интервала $(x_{i-1},x_i)$, будет являться значением миноранты в точке ее миинмума на интервале $(x_{i-1},x_i), 1\leq i\leq k+p$.
@@
 Теория сходимости и различные модификации последовательных и параллельных алгоритмов детально изложены в монографиях \cite{Strongin2000,Sergeyev2017}. 
-Отметим, что описанная здесь асинхронная схема распалаллеливания в отличие от синхронных схем, использовавшихся ранее при решении ряда прикладных задач (see \cite{Kalyulin2017,Modorskii2016}), обеспечивает полную загрузку всех задействованных процессов при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска.
+Отметим, что описанная здесь асинхронная схема распараллеливания в отличие от синхронных схем, использовавшихся ранее при решении ряда прикладных задач (see \cite{Kalyulin2017,Modorskii2016}), обеспечивает полную загрузку всех задействованных процессов при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска.
 
@@
 \center{\includegraphics[width=0.9\linewidth]{fig2.JPG}}
-\caption{Схема параллельных вычислений в системе Globalizer \\ (1 -- MPI связь, 2 -- связь через потоки).}\label{fig:Impl}
+\caption{Реализованная схема параллельных вычислений (1 -- link via MPI, 2 -- link via streams).}\label{fig:Impl}
 \end{figure}
@@
 В нашей реализации алгоритма глобального поиска для организации параллельных вычислений используется MPI.
-Процесс с рангом 0 выполняет роль мастера.
-В рамках процесса реализован асинхронный алгоритм глобального поиска (см.
-раздел 3.2 шаги 1-5).
-Процессы с рангом отличным от 0 выполняют роль вычислителей (рабочих).
+Процесс с рангом 0 выполняет роль мастера, процессы с рангом отличным от 0 выполняют роль вычислителей (см.
+описание алгоритма в предыдущем разделе).
 
 Принцип взаимодействия процессов (рисунок \ref{fig:Impl}, связь 1) заключается в следующем.
-В рамках первого шага на мастере выполняется выбор $p$ точек для проведения испытаний (см.
-раздел 3.2).
+В рамках первого шага в процессе-мастере выполняется выбор $p$ точек для проведения испытаний.
 Выбранные точки проведения испытаний рассылаются вычислителям и мастер переходит в состояние ожидания приема сообщений.
@@
 Новая точка отправляется в виде сообщения на вычислитель.
-Если в GSA выполнено условие остановки, то вместо точки для испытания на вычислитель отправляется сообщение о завершении работы.
+Если в алгоритме выполнено условие остановки, то вместо точки для проведения испытания мастер отправляет на вычислитель сообщение о завершении работы.
 
-Каждый процесс, выполняющий роль вычислителя, при старте создает экземпляр задачи GSA, а затем переходит в состояние ожидания сообщений от мастер-процесса.
+Каждый процесс, выполняющий роль вычислителя, при старте создает экземпляр задачи, а затем переходит в состояние ожидания сообщений от мастер-процесса.
 При приеме сообщения содержащего точку испытания, выполняется вычислительный эксперимент.
@@
 После проведения эксперимента сам процесс продолжает ожидание новых сообщений.
-При приеме сообщения о завершении вычислений выполняется высвобождение ресурсов, используемых задачей GSA, и завершение основного процесса.
+При приеме сообщения о завершении вычислений выполняется высвобождение ресурсов, используемых задачей, и завершение основного процесса.
 
@@
 
-При проведении вычислительных экспериментов (см.
-раздел 5) задача химической кинетики была реализована в рамках MATLAB.
+При проведении вычислительных экспериментов (see next section) решение прямой задачи химической кинетики выполнялось с помощью MATLAB.
 В исходном варианте задачи каждый набор параметров требовал запуск процесса MATLAB, передачу точки проведения испытаний, а также получение результатов вычислительных экспериментов.
-Для сокращения накладных расходов на организацию вычислений были выполнены ряд оптимизаций.
+Для сокращения накладных расходов на организацию вычислений был выполнены ряд дополнительных настроек.
 Во-первых, исходный код для вычислительного эксперимента был скомпилирован в исполняемое приложение.
-Во-вторых, был реализован подход, при котором MATLAB код выполнял ожидание очередной точки испытаний через входной поток, а через выходной поток осуществлял возврат результатов вычислительных экспериментов.
-Подобный подход позволил в каждом вычислителе один раз запустить процесс с откомпилированным MATLAB кодом, перехватить входные и выходные потоки, а затем выполнять вычисления без ожидания запуска новых процессов.
+Во-вторых, был реализован подход, при котором MATLAB получал очередную точку для проведения испытания через input stream, а через output stream осуществлял возврат результатов вычислительных экспериментов.
+Подобный подход позволил в каждом вычислителе один раз запустить процесс с откомпилированным кодом MATLAB, перехватить input and output streams, а затем выполнять вычисления без ожидания запуска новых процессов.
 

### 2022-08-19 2022/PAVT Special Issue/template.tex
reorder UNN part

@@
 
-%Стыковочный абзац - его надо переписать
-В сложных задачах химической кинетики найти кинетические константы реакций аналитически не представляется возможным, поэтому возникает необходимость в применеии численных методов для их поиска (см., например, [ссылки]).
+%Часть ННГУ - начало
+
+В сложных задачах химической кинетики найти кинетические константы реакций аналитически не представляется возможным, поэтому возникает необходимость в применении численных методов для их поиска (see, e.g., [ссылки]).
 В этом случае критерий качества (objective function) задается в виде некоторого алгоритма, который включает в себя этап численного моделирования и, как следствие, является трудоемким.
@@
 
-%Часть ННГУ - начало
 Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) существенно отличаются от методов локальной оптимизации, т.к.
@@
 В силу быстроты работы и малых требований к вычислительным ресурсам (в частности, памяти) метаэвристические алгоритмы получили широкое распространение при решении практических задач.
-Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может распологаться далеко от глобального оптимума \cite{Kvasov2018}.
+Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может располагаться далеко от глобального оптимума \cite{Kvasov2018}.
 
@@
 Одним из естественных допущений о задаче является предположение о том, что изменение значений целевой функции ограничено изменением ее аргумента.
-В таком случае, как известно, функции называют липшицевыми, а саму задачу --- Lipschitz global optimization problem.
+В таком случае говорят, что функция удовлетворяет Lipschitz condition, а саму задачу называют Lipschitz global optimization problem.
 Центральным направлением развития современных методов липшицевой оптимизации является разработка параллельных алгоритмов.
-При этом предполагается, что значения целевой функции вычисляются в результате решения какой-либо другой вычислительной задачи и требуют проведения затратного вычислительного эксперимента.
+При этом предполагается, что значения целевой функции определяются в результате решения какой-либо другой задачи и требуют проведения затратного вычислительного эксперимента.
 
@@
 Основная часть статьи имеет следующую структуру. 
+Описание математической модели исследуемой химической реакции приведено в Section 2.
+Формальная постановка задачи липшицевой глобальной оптимизации и алгоритм ее решения рассмотрены в Section 3.
+Здесь же изложены вопросы, связанные с техническими аспектами реализации параллельной реализации алгоритма.
+Результаты численного решения обратной задачи химической кинетики с использованием разработанного параллельного алгоритма обсуждаются в Section 4.
+Последний раздел содержит заключение и планы будущей работы.
+
 %Часть ННГУ - конец
-
-%Описание математической модели исследуемой химической реакции приведено в Section 2.
-Формальная постановка задачи липшицевой глобальной оптимизации приводятся в Section 3.
-В Section 4 кратко изложена схема разрабатываемого авторами асинхронного параллельного алгоритма для решения задач указанного класса.
-Результаты численного решения обратной задачи химической кинетики обсуждаются в Section 5.
 
@@
 %Часть ННГУ - начало
-%поправить название
 \section{Parallel Algorithm for Solving Global Optimization Problems }\label{Sec_GSA}
@@
 
-\subsection{Figures, Tables and Schemes}
-
-All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, Table~\ref{tab2}, etc.
-
-\begin{figure}[H]
-\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
-\caption{This is a figure.
-Schemes follow the same formatting.
-If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
-(\textbf{b}) Description of what is contained in the second panel.
-Figures should be placed in the main text near to the first time they are cited.
-A caption on a single line should be centered.\label{fig1}}
-\end{figure}   
-\unskip
-
-\begin{table}[H] 
-\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
-\newcolumntype{C}{>{\centering\arraybackslash}X}
-\begin{tabularx}{\textwidth}{CCC}
-\toprule
-\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-\midrule
-Entry 1		& Data			& Data\\
-Entry 2		& Data			& Data\\
-\bottomrule
-\end{tabularx}
-\end{table}
-\unskip
-
-\begin{table}[H]
-\caption{This is a wide table.\label{tab2}}
-	\begin{adjustwidth}{-\extralength}{0cm}
-		\newcolumntype{C}{>{\centering\arraybackslash}X}
-		\begin{tabularx}{\fulllength}{CCCC}
-			\toprule
-			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
-			\midrule
-			Entry 1		& Data			& Data			& Data\\
-			Entry 2		& Data			& Data			& Data \textsuperscript{1}\\
-			\bottomrule
-		\end{tabularx}
-	\end{adjustwidth}
-	\noindent{\footnotesize{\textsuperscript{1} This is a table footnote.}}
-\end{table}
-
-\subsection{Formatting of Mathematical Components}
-
-This is the example 1 of equation:
-\begin{linenomath}
-\begin{equation}
-a = 1,
-\end{equation}
-\end{linenomath}
-the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
-%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
-This ensures correct linenumbering.
-The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
-
-This is the example 2 of equation:
-\begin{adjustwidth}{-\extralength}{0cm}
-\begin{equation}
-a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
-\end{equation}
-\end{adjustwidth}
-
-% Example of a figure that spans the whole page width. The same concept works for tables, too.
-\begin{figure}[H]
-\begin{adjustwidth}{-\extralength}{0cm}
-\centering
-\includegraphics[width=13.5cm]{Definitions/logo-mdpi}
-\end{adjustwidth}
-\caption{This is a wide figure.\label{fig2}}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-08-19 2022/PAVT Special Issue/template.tex
before translation of UNN part

@@
 % Дежурное название: Использование ...алгоритма для определения кинетических параметров процесса каталитической изомеризации пентан-гексановой фракции
+% Еще один вариант: О решении задачи определения Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction с помощью Parallel Global Search Algorithm
+% На английском: About Solving the Problem of Finding Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction Using Parallel Global Search Algorithm
 
@@
 
-%Нужен комментарий про практическую реализацию алгоритма - используемые технологии, основные моменты реализации и т.п.
-
-%Асинхронный алгоритм глобального поиска был реализован с помощью технологии MPI, язык разработки --- С++.
-
 \subsection{Особенности программной реализации параллельного алгоритма}
@@
 Подобный подход позволил в каждом вычислителе один раз запустить процесс с откомпилированным кодом MATLAB, перехватить input and output streams, а затем выполнять вычисления без ожидания запуска новых процессов.
+
+\section{Numerical Experiments}
+
+Асинхронный параллельный алгоритм глобального поиска был реализован на C++, для сборки программы использовался Intel MPI и компилятор Intel C++ версии 2017.0.1.
+Целевая функция решаемой задачи была реализована в MATLAB 2019b.
+С целью проверки стабильности разработанной параллельной программы были проведены ее запуски на трех разных суперкомпьютерах: Lomonosov-2 (Moscow State University), Lobachevsky (University of Nizhni Novgorod), and MVS-10P (Joint Supercomputer Center of RAS).
+Основная часть экспериментов проводилась на supercomputer Lomonosov-2 с вычислительными узлами, обладающими следующими характеристиками: центральный процессор Intel Haswell-EP E5-2697v3 (частота процессора 2.6 GHz, 14 cores), объем оперативной памяти 64 GB, операционная система CentOS 7.
+
+Для решения оптимизационной задачи использовались комбинация параллельного алгоритма глобального поиска и метода локальной оптимизации.
+На первом этапе поиска использовался параллельный асинхронный алгоритм, описанный в Section \ref{Sec_GSA}.
+При запуске был задействован 101 процесс, из которых один процесс был мастером, а остальные 100 --- рабочими (вычислителями).
+Значения параметров метода были установлены следующим образом: максимальное число испытаний $K_{max}=3000$, точность поиска $\varepsilon=10^{-3}$, параметр надежности $r=3$, параметр точности построения развертки $m=12$.
+На втором этапе найденного решение уточнялось с помощью pattern search method с точностью $\varepsilon=10^{-4}$ и ограничением на число испытаний $K_{max}=500$.
+Общее время решения составило 8109 сек., найдена точка оптимума со значением целевой функции $\phi^* = 0.002115$.
 

### 2022-08-19 2022/PAVT Special Issue/template.tex
add one more reference

@@
 Значения параметров метода были установлены следующим образом: максимальное число испытаний $K_{max}=3000$, точность поиска $\varepsilon=10^{-3}$, параметр надежности $r=3$, параметр точности построения развертки $m=12$.
-На втором этапе найденного решение уточнялось с помощью pattern search method с точностью $\varepsilon=10^{-4}$ и ограничением на число испытаний $K_{max}=500$.
+На втором этапе найденного решение уточнялось с помощью pattern search method (see, e.g., \cite{Torczon1997}) с точностью $\varepsilon=10^{-4}$ и ограничением на число испытаний $K_{max}=500$.
 Общее время решения составило 8109 сек., найдена точка оптимума со значением целевой функции $\phi^* = 0.002115$.
@@
 %=====================================
-\bibliography{bibliography}
+ё\bibliography{bibliography}
 

### 2022-08-29 2022/PAVT Special Issue/template.tex
UNN part translated

--- 2022-08-19 2022/PAVT Special Issue/template.tex
@@
 
-В сложных задачах химической кинетики найти кинетические константы реакций аналитически не представляется возможным, поэтому возникает необходимость в применении численных методов для их поиска (see, e.g., [ссылки]).
-В этом случае критерий качества (objective function) задается в виде некоторого алгоритма, который включает в себя этап численного моделирования и, как следствие, является трудоемким.
-Кроме того, в обратных задачах химической кинетики целевая функция может быть существенно многоэкстремальной, т.е.
-иметь много локальных экстремумов наряду с глобальным.
+In the complex problems of chemical kinetics, kinetic constants of reactions cannot be found analytically, they can only be found using numerical methods (see, for example, [references]).
+In this case, the quality criterion (objective function) is set in the form of a certain algorithm, which includes the stage of numerical modeling, which, in turn, makes it compute-intensive.
+Furthermore, in the inverse problems of chemical kinetics, the objective function can be substantially multiextremal, i.e., having many local extremes along the global one.
 
-Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) существенно отличаются от методов локальной оптимизации, т.к.
-требуют исследования всей search domain (see, e.g., \cite{PaulaviciusZilinskas2014,Sergeyev2017}).
-Указанные методы можно условно разделить на два класса: метаэвристические и детерминированные.
-Метаэвристические алгоритмы, как правило, основаны на имитации процессов, протекающих в живой природе.
-Примерами метаэвристических алгоритмов являются simulated annealing, evolution and genetic algorithms и т.д. (see, e.g., \cite{Battiti2009,Eiben2015}).
-Указанные алгоритмы входят во многие библиотеки программных средств для решения задач глобальной оптимизации (например, MATLAB Global Optimization Toolbox, Scipy.Optimize и др. ).
-В силу быстроты работы и малых требований к вычислительным ресурсам (в частности, памяти) метаэвристические алгоритмы получили широкое распространение при решении практических задач.
-Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может располагаться далеко от глобального оптимума \cite{Kvasov2018}.
+Numerical methods for solving such multiextremal problems (global optimization methods) differ significantly from local optimization methods since they require the study of the entire search domain (see, for example, \cite{PaulaviciusZilinskas2014,Sergeyev2017}).
+These methods can be divided into two classes: meta-heuristic and deterministic.
+Metaheuristic algorithms are usually based on simulations of processes occurring in the wild.
+Examples of metaheuristic algorithms are simulated annealing, evolution and genetic algorithms, etc. (see, for example, \cite{Battiti2009,Eiben2015}).
+These algorithms are included in many software libraries for solving global optimization problems (for example, MATLAB Global Optimization Toolbox, Scipy.Optimize, etc.).
+Due to the operation speed and low requirements for computing resources (in particular, memory), metaheuristic algorithms have become widespread for solving practical problems.
+However, the solution to the problem found by such an algorithm is, in general, local and can be located far from the global optimum \cite{Kvasov2018}.
 
-Возможность построения детерминированных методов глобального поиска, отличающихся от brute force algorithm и метаэвристических алгоритмов, связана с наличием и учетом  априорных предположений о свойствах функций задачи.
-Одним из естественных допущений о задаче является предположение о том, что изменение значений целевой функции ограничено изменением ее аргумента.
-В таком случае говорят, что функция удовлетворяет Lipschitz condition, а саму задачу называют Lipschitz global optimization problem.
-Центральным направлением развития современных методов липшицевой оптимизации является разработка параллельных алгоритмов.
-При этом предполагается, что значения целевой функции определяются в результате решения какой-либо другой задачи и требуют проведения затратного вычислительного эксперимента.
+It is also possible to construct deterministic methods of global search, which are different from brute force algorithms and metaheuristic algorithms.
+Such methods are associated with the presence and consideration of a priori assumptions about the problem properties.
+One of the natural assumptions about the problem is that changing the values of an objective function is limited by changes of its argument.
+In this case, the function satisfies the Lipschitz condition, and the problem itself is called the Lipschitz global optimization problem.
+The central direction of the development of modern methods of Lipschitz optimization is the development of parallel algorithms.
+It is assumed that the values of the objective function are determined as a result of solving some other task and require a costly computational experiment.
 
-В данной статье отражены результаты применения параллельных методов липшицевой оптимизации, развиваемых в работах [], к решению inverse problems of chemical kinetics. 
-Основная часть статьи имеет следующую структуру. 
-Описание математической модели исследуемой химической реакции приведено в Section 2.
-Формальная постановка задачи липшицевой глобальной оптимизации и алгоритм ее решения рассмотрены в Section 3.
-Здесь же изложены вопросы, связанные с техническими аспектами реализации параллельной реализации алгоритма.
-Результаты численного решения обратной задачи химической кинетики с использованием разработанного параллельного алгоритма обсуждаются в Section 4.
-Последний раздел содержит заключение и планы будущей работы.
+This paper reflects the results of applying parallel methods of Lipschitz optimization developed in [refernces], to the solution of inverse problems of chemical kinetics.  
+The main part of the paper is structured as follows. 
+Section 2 provides a description of the mathematical model of the chemical reaction under study.
+Section 3 discusses the standard formulation of the Lipschitz global optimization problem and the algorithm for solving it.
+It also outlines the issues related to the technical aspects of the implementation of the parallel algorithm.
+The results of the numerical solution of the inverse problem of chemical kinetics using the developed parallel algorithm are discussed in Section 4.
+The last section contains the conclusion and plans for future work.
 
@@
 
-\subsection{Постановка задачи глобальной оптимизации}
+\subsection{Statement of the global optimization problem}
 
-Как уже говорилось выше, с математической точки зрения the inverse problem of chemical kinetics может рассматриваться как Lipschitz global optimization problem. 
-В общем виде задача указанного класса может быть сформулирована следующим образом:
+As mentioned above, from a mathematical point of view, the inverse problem of chemical kinetics can be considered a Lipschitz global optimization problem. 
+In general, the problem of the specified class can be formulated as follows:
 \begin{gather}
@@
 \end{gather}
-где $a,b \in R^N$ есть заданные границы области поиска, а целевая функция $\phi(y)$ удовлетворяет the Lipschitz condition
+where $a,b \in R^N$ are boundaries of the search domain, and the objective function $\phi(y)$ satisfies the Lipschitz condition
 \begin{equation}\label{Lip}
@@
 
-Функция $\phi(y)$ предполагается многоэкстремальной и заданной в виде некоторого вычислительного алгоритма, на вход которого подается вектор параметров, а выходом является  соответствующее значение функции.
-Подобные функции часто называются функциями вида ``black box''.
-Также предполагается, что каждое вычисление значение функции в точке области поиска (называемое в дальнейшем также trial) является трудоемкой операцией и требует значительных вычислительных ресурсов.
-Как отмечалось во введении, такая постановка задачи полностью соответствует the inverse problem of chemical kinetics.
+The function $\phi(y)$ is assumed to be multiextremal and is given in the form of some computational algorithm.
+The input of this algorithm is supplied with a parameter vector, and the output is the corresponding value of the function.
+Such functions are often referred to as ``black-box'' functions.
+It is also assumed that each calculation of the function value at the search domain point (hereinafter also referred to as trial) is a laborious operation and requires significant computational resources.
+As noted in the introduction, this statement of the problem fully corresponds to the inverse problem of chemical kinetics.
 
-The Lipschitz condition (\ref{Lip}) может быть использовано для построения оценок глобального минимума функции в области поиска, на основе которых можно конструировать алгоритмы глобальной оптмизации и доказывать условия их сходимости (see, for example, \cite{Strongin2000}).
+The Lipschitz condition (\ref{Lip}) can be used to construct estimates of the function global minimum, which can serve as a foundation for designing global optimization algorithms and proving their convergence conditions (see, for example, \cite{Strongin2000}).
 
-Одной из основных трудностей при решении Lipschitz global optimization problems является так называемое ``проклятие размерности'' --- экспоненциальный рост числа trials, требующегося для решения задачи с заданной точностью, при увеличении размерности задачи.
-Сгладить данный эффект при сохранении точности решения возможно за счет построения существенно неравномерных покрытий области поиска точками испытаний.
+One of the main difficulties in solving Lipschitz global optimization problems is the so-called ``curse of dimensionality'' --- the exponential increase in the number of trials required to solve a problem with a given accuracy, accompanied by the increasing dimensionality of the problem.
+It is possible to smooth this effect while maintaining the accuracy of the solution by constructing significantly uneven coverages of the search domain with trial points.
 
-Известными примерами таких методов являются non-uniform space covering method \cite{Evtushenko2013}, simplicial partitions method \cite{PaulaviciusZilinskas2014}, diagonal partitions method \cite{Sergeyev2017}.
-Указанные методы основаны на адаптивном разбиении области поиска на конечное число подобластей на каждой итерации.
-Для указанных подобластей оценивается (с использованием накопленной информации о целевой функции) необходимость дальнейшего подразбиения для более детального исследования. 
- Принципиально иным подходом к решению многомерной задачи (\ref{problemN}) является ее редукция к одной или нескольким одномерным задачам с последующим применением эффективных методов одномерной оптмизации.
-Такая редукция может быть осуществлена, например, при помощи nested optimization scheme \cite{Grishagin2018} или при помощи Peano-Hilbert curves \cite{Barkalov2018}.
-Последний из указанных подходов и используется в настоящей работе.
+Known examples of such methods are the non-uniform space covering method \cite{Evtushenko2013}, the simplicial partitions method \cite{PaulaviciusZilinskas2014}, the diagonal partitions method \cite{Sergeyev2017}.
+These techniques are based on adaptive partitioning of the search domain into a finite number of subdomains at each iteration.
+Then the need for further subdivision for a more detailed study of these subdomains is estimated (using the accumulated information about the objective function). 
+A fundamentally different approach to solving a multidimensional problem (\ref{problemN}) is its reduction to one or more one-dimensional problems followed by the use of efficient methods of one-dimensional optimization.
+Such reduction may be accomplished, for example, by a nested optimization scheme \cite{Grishagin2018} or by Peano-Hilbert curves \cite{Barkalov2018}.
+The latter approach is used in this work.
 
-Используя непрерывное однозначное отображение (Peano-Hilbert curve) $y(x)$ отрезка $[0,1]$ вещественной оси на гиперинтервал $D$ из (\ref{D}) можно свести многомерную задачу (\ref{problemN}) к одномерной задаче
+Using a continuous single-valued mapping (Peano-Hilbert curve) $y(x)$ of the real axis segment $[0,1]$ onto hyperinterval $D$ from (\ref{D}), a multidimensional problem (\ref{problemN}) can be reduced to a one-dimensional one
 \[
@@
 \]
-Известно \cite{Strongin2000,Sergeyev2013}, что редуцированная одномерная функция $\phi(y(x))$ уже не будет липшицевой, но будет удовлетворять H{\"o}lder condition
+It is known \cite{Strongin2000, Sergeyev2013} that the reduced one-dimensional function $\phi(y(x))$ will no longer be Lipschitzian, but will satisfy the H{\"o}lder condition
 \[
-\left|\phi(y(x_1))-\phi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}
+\left|\phi(y(x_1))-\phi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N},
 \]
-где H{\"o}lder constant $H$ связана с Lipschitz constant $L$ соотношением $ H=2 L \sqrt{N+3}$.
+where the H{\"o}lder constant $H$ is associated with the Lipschitz constant $L$ by the relationship $ H=2 L \sqrt{N+3}$.
 
-Конкретные способы численного построения подобного рода оботражений (называемых evolvents) рассмотрены в \cite{Strongin2000,Sergeyev2013}.
-Здесь мы отметим, что evolvent будет аппроксимировать теоретическую Peano-Hilbert curve с точностью $2^{-m}$, где $m$ --- параметр, задающий точность аппроксимации.
-Примеры evolvent для разных размерностей приведены на Fig.~\ref{fig:Peano}.
+Specific methods of numerical construction of this kind of mappings (called evolvents) are discussed in \cite{Strongin2000,Sergeyev2013}.
+Here we note that evolvent will approximate the theoretical Peano-Hilbert curve with an accuracy of $2^{-m}$, where $m$ is the parameter that sets the approximation accuracy.
+Examples of evolvent for different dimensions are shown in Fig.~\ref{fig:Peano}.
 
@@
 
-Таким образом, в рамках данного подхода search trial в некотрой точке $x^k\in[0,1]$ будет включать в себя сначала the construction of the image $y^k=y(x^k)$, и лишь затем вычисление значения функции $ z^k = \phi(y^k)$.
-Тройку значений $\{x^k,y^k,z^k\}$ мы будем называть outcome проведенного испытания.
+Thus, under this approach, the search trial at some point $x^k\in[0,1]$ will include the construction of the image $y^k=y(x^k)$ first, and only then the calculation of the value of the function $ z^k = \phi(y^k)$.
+We will call the three values $\{x^k,y ^k,z^k\}$ the trial result.
 
 %поправить название
-\subsection{Асинхронный параллельный алгоритм поиска глобального минимума функции}
+\subsection{Asynchronous parallel algorithm used to determine the global minimum of a function}
 
-Используемый в данном исследовании параллельный алгоритм основан на информационно-статистическом подходе к разработке методов глобальной оптимизации, теоретические основы которого изложены в \cite{Strongin2000}.
+The parallel algorithm used in this study is based on the information and statistical approach to the development of global optimization methods, the theoretical foundations of which are set out in \cite{Strongin2000}.
 
-Нами была реализована асинхнонная схема распараллеливания вида ``master/worker''.
-В процессе-мастере выполняется алгоритм глобального поиска, в котором проводится накопление поисковой информации, оценка на ее основе константы Липшица для целевой функции, вычисление точек новых испытаний и распределение их по рабочим процессам.
-Процессы-рабочие получают от мастера точки, проводят в них новые испытания и отсылают мастеру результаты испытаний.
+We have implemented an asynchronous paralleling scheme of the ``master/worker'' type.
+The master process performs a global search algorithm, in which the search information is accumulated.
+Then the Lipschitz constants for the objective function are estimated on its basis, and the points of new trials are calculated and distributed over the work processes.
+Worker-processes receive points from the master, carry out new trials in them, and send the results of the trials to the master.
 
-Будем считать, что на каждой итерации процесс-мастер вычисляет одну точку нового испытания и передает ее для проведения испытания в процесс-рабочий.
-При этом проведение испытания рабочим является значительно более трудоемкой операцией, чем выбор новой точки мастером, что исключает простой процессов-рабочих.
-В данном случае (в отличие от синхронных параллельных алгоритмов) общее число выполненных испытаний каждым процессом-рабочим будет зависеть от трудоемкости проведения конкретного испытания и не может быть оценено заранее.
+Let us assume that at each iteration, the master-process calculates one point of the new trial and transfers it for testing to the worker-process.
+At the same time, a trial performed by a worker is a much more labor-intensive operation than selection of a new point by a master, which eliminates the downtime of worker-processes.
+In this case (unlike synchronous parallel algorithms), the total number of trials performed by each worker-process  will depend on the labor intensity of the specific trial and cannot be estimated in advance.
 
-При описании параллельного алгоритма будем предполагать, что есть один процесс-мастер и $p$ рабочих процессов.
-Верхним индексом будем обозначать порядковый номер проведенного испытания, а нижний индекс будем использовать для нумерации испытаний в порядке возрастания координаты.
+When describing a parallel algorithm, let us assume that there is one process master and $p$ workers.
+The upper index will indicate the serial number of the trial performed, and the lower index will be used to number the trials in ascending order of the coordinate.
 
-Начальная итерация проводится по особому правилу.
-Процесс-мастер (будем считать, что это процесс No 0) инициирует параллельное проведение $p$ испытаний в $p$ различных точках области поиска, две из которых являются граничными, а остальные --- произвольными внутренними точками (например, распределенными равномерно).
-Таким оборазом, начальные испытания проводятся в точках $\{y(x^1), y(x^2), ...,y(x^p)\}$, где $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.
+The initial iteration is carried out according to a special rule.
+The master process (let us assume that this is the process No 0) initiates  $p$ parallel trials at $p$ different points of the search domain, two of which are boundary points, and the rest are arbitrary internal points (for example, uniformly distributed).
+Thus, the initial trials are carried out at the points $\{y(x^1), y(x^2), ...,y(x^p)\}$, where $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.
 
-Предположим теперь, что процессом-мастером получены результаты $k$ испытаний, и в текущий момент времени процессы-рабочие проводят испытания в точках $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.
+Let us suppose that the master-process has obtained the results of $k$ trials, and at the current time the worker-processes are conducting trials at the points
+$\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.
 
-Пусть один из процессов (не ограничивая общности, можно считать, что это процесс No 1), в некоторый момент времени завершил проведение испытания в точке $y(x^{k+1})$.
-Остальные процессы находятся в стадии проведения своих испытаний, т.е. в точках множества 
-\[
-I_k = \left\{ x^{k+2},...,x^{k+p} \right\}
-\]
-испытания уже начались, но еще не были завершены.
+Let one of the processes (without limiting the commonality, we can assume that this is the process No 1), at some point in time complete the trial at the point $y(x^{k+1})$.
+The remaining processes are at the stage of conducting their trials, i.e. at these points multiple trials have already started, but have not yet been completed.
 
-Тогда процесс No 1 пересылает процессу-мастеру полученные им результаты испытания, т.е. тройку  $\{x^{k+1},y^{k+1},z^{k+1}\}$.
-Мастер сохраняет полученные результаты в своей информационной базе и выбирает для процесса No 1 точку нового испытания $x^{k+p+1}$ в соответствии с правилами, описанными ниже.
+Then the No 1 process forwards to the master process the trial results, i.e. the three values $\{x^{k+1},y^{k+1},z^{k+1}\}$.
+The master saves the results in its information base and selects a new trial point $x^{k+p+1}$ for the process No 1 in accordance with the rules described below.
 
-Step 1. Перенумеровать нижним идексом в порядке возрастания координаты точки множества
+Step 1. Renumber he point of the set 
 \[
@@
 \]
-которое содержит все точки, в которых либо проведены, либо проводятся испытания, т.е.
+which contains all the points at which the trials either have been carried out or are being carried out by the lower index in ascending order, i.e.
 \[
@@
 \]
-Step 2. Вычислить значения 
+Step 2. Calculate values
 \[
@@
 \]
-где $z_i=\phi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$.
-Значения $z_i$ в точках $x_i \in I_k$ являются неопределенными, т.к.
-испытания в точках $x_i \in I_k$ еще не завершены.
-Если значение $M$ получилось равным 0, то присвоить $M=1$.
+where $z_i=\phi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$.
+The $z_i$ values at $x_i \in I_k$ are undefined because the trials at $x_i \in I_k$ are not yet complete.
+If the value $M$ is 0, assign $M=1$.
 
-Step 3. Для каждого интервала $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, вычислить значение 
+Step 3. For each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, calculate the value 
 \[
@@
 \]
-называемое характеристикой данного интервала. Здесь $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.
+called the characteristic of this interval. Here $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.
 
-Step 4. Выбрать интервал $[x_{t-1},x_t]$ с наибольшим значением характеристики, т.е.
+Step 4. Select the interval $[x_{t-1},x_t]$ with the highest characteristic value, i.e.
 \[
@@
 
-Step 5. Вычислить точку $x^{k+p+1} \in (x_{t-1},x_t)$ в соответствии с формулой
+Step 5. Calculate the point $x^{k+p+1} \in (x_{t-1},x_t)$ according to the formula
 \[
@@
 
-Сразу после вычисления точки $x^{k+p+1}$ очередного испытания процесс-мастер добавляет ее в множество $I_k$ и пересылает процессу-рабочему, который инициирует проведение испытания в ней.
+Immediately after calculating the point $x^{k+p+1}$ of the next trial, the master-process adds it to the set $I_k$ and forwards it to the worker-process, which initiates the trial at this point.
 
-Процесс-мастер завершает работу алгоритма, если выполняется либо условие $\Delta_{t}<\varepsilon$ ($t$ --- номер интервала c максимальной характеристикой из Правила 4), либо условие $k>K_{max}$ ($k$ --- число выполненных испытаний).
-Вещественное число $\varepsilon>0$ и целое число $K_{max}>0$ являются параметрами алгоритма и соответствуют точности поиска решения и максимальному числу испытаний.
+The master process terminates the algorithm if either the condition $\Delta_{t}<\varepsilon$ ($t$ is number of the interval with the maximum characteristic from Step 4) or the condition $k>K_{max}$ ($k$ is the number of trials performed) is met.
+The real number $\varepsilon>0$ and the integer $K_{max}>0$ are the algorithm parameters which correspond to the accuracy of the solution search and the maximum number of trials.
 
-Последовательный алгоритм глобального поиска, на основе которого был реализован предложенный асинхронный алгоритм, допускает следующую интерпретацию \cite{Molinaro2001}.
-Используя точки ранее проведенных испытаний можно построить миноранту целевой функции, для которой значение $-R(i)$, где $R(i)$ --- значение характеристики интервала $(x_{i-1},x_i)$, будет являться значением миноранты в точке ее миинмума на интервале $(x_{i-1},x_i), 1\leq i\leq k+p$.
-В соответствии с правилом 4 алгоритма, новое испытание проводится в интервале с наименьшим значением миноранты.
-А в соответствии с правилом 5 точка нового испытания совпадает с точкой минимума миноранты.
+The sequential global search algorithm, on the basis of which the proposed asynchronous algorithm was implemented, allows the following interpretation \cite{Molinaro2001}.
+Using the points of previously performed trials, it is possible to construct a minorant of the objective function, for which the value $-R(i)$, where $R(i)$ is the value of the characteristic of the interval $(x_{i-1}, x_i)$, will be the value of the minorant at the point of its minimum at the interval $(x_{i-1},x_i), 1\leq i\leq k+p$.
+In accordance with Step 4 of the algorithm, a new trial is carried out in the interval with the lowest value of the minorant.
+And in accordance with Step 5, the point of the new trial coincides with the minorant minimum point.
 
-Теория сходимости и различные модификации последовательных и параллельных алгоритмов детально изложены в монографиях \cite{Strongin2000,Sergeyev2017}. 
-Отметим, что описанная здесь асинхронная схема распараллеливания в отличие от синхронных схем, использовавшихся ранее при решении ряда прикладных задач (see \cite{Kalyulin2017,Modorskii2016}), обеспечивает полную загрузку всех задействованных процессов при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска.
+The theory of convergence and various modifications of serial and parallel algorithms are detailed in the monographs \cite{Strongin2000,Sergeyev2017}. 
+It should be noted that the asynchronous parallelization scheme described here, in contrast to the synchronous schemes used earlier in solving a number of application problems (see \cite{Kalyulin2017, Modorskii2016}), provides a full load of all the processes involved in solving problems with different labor intensity of trials at different points in the search domain.
 
-%Нужен комментарий про практическую реализацию алгоритма - используемые технологии, основные моменты реализации и т.п.
-
-%Асинхронный алгоритм глобального поиска был реализован с помощью технологии MPI, язык разработки --- С++.
-
-\subsection{Особенности программной реализации параллельного алгоритма}
+\subsection{Features of the software implementation of the parallel algorithm}
 
@@
 \center{\includegraphics[width=0.9\linewidth]{fig2.JPG}}
-\caption{Реализованная схема параллельных вычислений (1 -- link via MPI, 2 -- link via streams).}\label{fig:Impl}
+\caption{Implemented scheme of parallel computations (1 -- link via MPI, 2 -- link via streams).}\label{fig:Impl}
 \end{figure}
 
-В нашей реализации алгоритма глобального поиска для организации параллельных вычислений используется MPI.
-Процесс с рангом 0 выполняет роль мастера, процессы с рангом отличным от 0 выполняют роль вычислителей (см.
-описание алгоритма в предыдущем разделе).
+In our implementation of the global search algorithm, MPI is used to organize parallel calculations.
+A process with the rank 0 acts as a master, processes with a rank other than 0 act as calculators (workers), see the description of the algorithm in the previous section.
 
-Принцип взаимодействия процессов (рисунок \ref{fig:Impl}, связь 1) заключается в следующем.
-В рамках первого шага в процессе-мастере выполняется выбор $p$ точек для проведения испытаний.
-Выбранные точки проведения испытаний рассылаются вычислителям и мастер переходит в состояние ожидания приема сообщений.
-Сообщения от вычислителей содержат outcome проведенного испытания.
-Полученная информация используется для выбора точки очередного испытания.
-Новая точка отправляется в виде сообщения на вычислитель.
-Если в алгоритме выполнено условие остановки, то вместо точки для проведения испытания мастер отправляет на вычислитель сообщение о завершении работы.
+The principle of process interaction (Figure \ref{fig:Impl}, link 1) is as follows.
+The first step in the master-process is to select $p$ points for trials.
+The selected trial points are sent to the calculators and the master enters the stand-by state for receiving messages.
+Messages from the calculators contain the trial result.
+The obtained information is used to select the point of the next trial.
+The new point is sent as a message to the calculator.
+If the stopping condition is met in the algorithm, instead of the trial point the master sends a shutdown message to the calculator.
 
-Каждый процесс, выполняющий роль вычислителя, при старте создает экземпляр задачи, а затем переходит в состояние ожидания сообщений от мастер-процесса.
-При приеме сообщения содержащего точку испытания, выполняется вычислительный эксперимент.
-Результаты эксперимента отправляются мастеру.
-После проведения эксперимента сам процесс продолжает ожидание новых сообщений.
-При приеме сообщения о завершении вычислений выполняется высвобождение ресурсов, используемых задачей, и завершение основного процесса.
+Each process that performs the role of a calculator creates an instance of the task when it starts, and then goes into a stand-by mode for messages from the master-process.
+When a message containing a trial point is received, a computational experiment is performed.
+The results of the experiment are sent to the master.
+After the experiment, the process itself continues to wait for new messages.
+When a shutdown message is received, the resources used by the task are released and the main process is terminated.
 
-При реализации асинхронной схемы мы предполагали, что время начала вычисления значений функции критерия на разных процессах отличается.
-Наша реализация учитывает тот факт, что точка испытания, а также результаты вычислений представляют незначительный объем данных, который относительно быстро можно отправить по сети.
-Как результат, для взаимодействия мастера и вычислителей (рисунок \ref{fig:Impl}, связи 1) использовались базовые операции пересылки сообщений MPI типа точка-точка.
+When implementing the asynchronous scheme, we assumed that the start time of calculating the objective function values is different for each particular process.
+Our implementation takes into account the fact that the trial point as well as the calculation results represent a small amount of data that can be sent relatively quickly over the network.
+As a result, basic point-to-point MPI message forwarding operations were used to interact between the master and the calculators (Figure \ref{fig:Impl}, link 1).
 
-При проведении вычислительных экспериментов (see next section) решение прямой задачи химической кинетики выполнялось с помощью MATLAB.
-В исходном варианте задачи каждый набор параметров требовал запуск процесса MATLAB, передачу точки проведения испытаний, а также получение результатов вычислительных экспериментов.
-Для сокращения накладных расходов на организацию вычислений был выполнены ряд дополнительных настроек.
-Во-первых, исходный код для вычислительного эксперимента был скомпилирован в исполняемое приложение.
-Во-вторых, был реализован подход, при котором MATLAB получал очередную точку для проведения испытания через input stream, а через output stream осуществлял возврат результатов вычислительных экспериментов.
-Подобный подход позволил в каждом вычислителе один раз запустить процесс с откомпилированным кодом MATLAB, перехватить input and output streams, а затем выполнять вычисления без ожидания запуска новых процессов.
+When conducting computational experiments (see next section), the solution of the direct problem of chemical kinetics was performed using MATLAB.
+In the initial version of the problem, each set of parameters required starting the MATLAB process, transferring the trial point, as well as obtaining the results of computational experiments.
+A number of additional settings have been made to reduce computing overhead.
+First, the source code for the computational experiment was compiled into an executable application.
+Second, the approach in which MATLAB receives a trial point through input stream, and returns the results of computational experiment through output stream was implemented.
+This approach allowed each calculator to run a process with compiled MATLAB code once, intercept input and output streams, and then perform calculations without waiting for new processes to start.
+
+\section{Numerical Experiments}
+
+The asynchronous parallel global search algorithm was implemented on C++; the software was built on Intel MPI and Intel C++ compiler, version 2017.0.1.
+The objective function of the problem to be solved was implemented in MATLAB 2019b.
+In order to check the stability of the developed parallel program, it was launched on three different supercomputers: Lomonosov-2 (Moscow State University), Lobachevsky (University of Nizhni Novgorod), and MVS-10P (Joint Supercomputer Center of the RAS).
+The main part of the experiments was carried out on the Lomonosov-2 supercomputer with the following computer nodes characteristics: Intel Haswell-EP E5-2697v3 CPU (processor frequency 2.6 GHz, 14 cores), 64 GB of RAM, CentOS 7 operating system.
+
+To solve the optimization problem, a combination of the parallel global search algorithm and a local optimization method was used.
+In the first stage of the search, the parallel asynchronous algorithm described in Section \ref{Sec_GSA} was used.
+101 processes were involved in the launch, of which one process was a master and the remaining 100 were workers (calculators).
+The values of the method parameters were set as follows: the maximum number of trials $K_{max}=3000$, the search accuracy $\varepsilon=10^{-3}$, the reliability parameter $r=3$, the evolvent accuracy parameter $m=12$.
+At the second stage, the solution was refined using the pattern search method with an accuracy of $\varepsilon=10^{-4}$ and a limit on the number of trials $K_{max}=500$.
+The total time for solving the problem was 8109 seconds, an optimum point was found with the objective function value $\phi^* = 0.002115$.
 

### 2022-09-04 2022/NumMethOneAPI/NMP_authorGuidelinesRu_2022.tex
Статья по OneAPI в "вычислительные методы и программирование". Первая версия (без доп. текста)

% Windows: протестировано для MikTeX+TeXworks в режиме pdfLatex.
%
% Linux:
% Для получения pdf используйте команду  pdflatex rfa_2021.tex.
% Команда latex rfa_2021.tex выведет сообщение об ошибке.
%
\NeedsTeXFormat{LaTeX2e}
\documentclass[10pt,a4paper]{book}
\usepackage{NumMet_2022}
%--- Здесь можно вставить необходимые стили ---
% \usepackage{...}
%
%--- Здесь можно добавить свои команды --------
% \newcommand{}
\newcommand{\ceq}{\mathrel{\vcenter{\hbox{:=}}}}
%----------------------------------------------
\selectlanguage{russian}
\begin{document}
%===========================  ШАПКА СТАТЬИ =================================
\Article{Решение трудоемких задач многомерной глобальной оптимизации с использованием набора инструментов Intel oneAPI
}
    {Solving time-consuming multidimensional global optimization problems with Intel OneApi tools}

\Abstract{В рамках данной статьи рассматривается решение серии сложно вычислимых задач глобальной оптимизации. В качестве целевой выбирается функция, удовлетворяющая условию Липшица с заранее не известной константой. Инструментом для реализации параллельного алгоритма выбран Intel oneAPI, который позволяет писать один код как для центрального процессора, так и для графического ускорителя. В решении многомерных задач применяется метод, построенный на идее редукции размерности при помощи кривой Пеано, которая непрерывно и однозначно отображает отрезок вещественной оси на гиперкуб.}
    {The paper considers the process of solving a series of time-consuming global optimization problems. We suppose that the objective function satisfies the Lipschitz condition with a priory unknown Lipschitz constant. Intel OneApi tools, that allows one to write the same code for both the central processor and the graphics accelerator, are chosen for implementing the parallel global optimization algorithm. To solve multidimensional problems, we use an approach based on the idea of dimensionality reduction using the Peano curve, which continuously maps an interval of the real axis onto an n-dimensional cube.}

\Keywords{глобальная оптимизация,
многоэкстремальные функции,
параллельные вычисления,
редукция размерности,
графические ускорители,
Intel OneAPI.}
        {global optimization,
multiextremal functions,
parallel computing,
reduction of dimensionality,
graphics accelerators,
Intel OneApi.}

\Acknowledgements{Работа выполнена при поддержке программы Центра компетенций OneAPI в ННГУ, Министерства науки и высшего образования РФ (проект № 0729-2020-0055) и научно-образовательного математического центра «Математика технологий будущего» (проект № 075-02-2021-1394).}
    {Работа выполнена при поддержке программы Центра компетенций OneAPI в ННГУ, Министерства науки и высшего образования РФ (проект № 0729-2020-0055) и научно-образовательного математического центра «Математика технологий будущего» (проект № 075-02-2021-1394).}

\Citation{К.А. Баркалов, И.Г. Лебедев, Я.В. Кольтюшкина
    Решение трудоемких задач многомерной глобальной оптимизации с использованием набора инструментов Intel oneAPI~//
    Вычислительные методы и программирование. 2022.
    \textbf{??}, \No~?. \pageref*{firstPage}--\pageref*{LastPage}.  
    doi 10.26089/NumMet.v??r???.}
    {K.A. Barkalov, I.G. Lebedev, Ya.V. Koltyushkina,
    Solving time-consuming multidimensional global optimization problems with Intel OneApi tools 
    for the journal ``Numerical methods and programming,''
    Numerical Methods and Programming. \textbf{??} (?), 
    \pageref*{firstPage}--\pageref*{LastPage} (2022).
    doi~10.26089/NumMet.v??r???.}

\UDC{<указывает автор>}
\DOI{10.26089/NumMet.v??r???}% заполняет редакция
\VOL{??}{?}
\YEAR{2021}
\Received{16 июня 2021 г.}{June 16, 2021}
\Accepted{13 октября 2021 г.}{October 13, 2021}

\Author{	К.А. Баркалов}{K.A. Barkalov}
\FullName{Баркалов Константин Александрович}{Barkalov Konstantin Aleksandrovich}
% Место работы 1
\Institution{Нижегородский государственный университет им. Н.И. Лобачевского}
{Nizhny Novgorod State University N.I. Lobachevsky}
\Address{пр. Гагарина, д. 23.}{Gagarin Ave., 23.}
\Postcode{603022}
\City{Нижний Новгород}{Nizhny Novgorod}
\CountryOfResidence{Российская Федерация}{Russia}
\AcademicDegree{
Доктор технических наук}{Ph.D.}
\Position{доцент}{docent}
\Orcid{0000-0001-5273-2471}
\Email{konstantinbarkalov@yandex.ru}

\Author{ И.Г. Лебедев}{ I.G. Lebedev}
\FullName{Лебедев Илья Генадьевич}{ Lebedev Ilya Genadievich}
% Место работы 1
\Institution{Нижегородский государственный университет им. Н.И. Лобачевского}
{Nizhny Novgorod State University N.I. Lobachevsky}
\Address{пр. Гагарина, д. 23.}{Gagarin Ave., 23.}
\Postcode{603022}
\City{Нижний Новгород}{Nizhny Novgorod}
\CountryOfResidence{Российская Федерация}{Russia}

\Position{заведующий лабораторией}{head of laboratory}
\Orcid{0000-0002-8736-0652}
\Email{ilya.lebedev@itmm.unn.ru}

\Author{ Я.В. Кольтюшкина}{ Ya.V. Koltyushkina}
\FullName{ Кольтюшкина Янина Вадимовна}{ Koltyushkina Yanina Vadimovna}
% Место работы 1
\Institution{Нижегородский государственный университет им. Н.И. Лобачевского}
{Nizhny Novgorod State University N.I. Lobachevsky}
\Address{пр. Гагарина, д. 23.}{Gagarin Ave., 23.}
\Postcode{603022}
\City{Нижний Новгород}{Nizhny Novgorod}
\CountryOfResidence{Российская Федерация}{Russia}
\Position{лаборант}{laboratory assistant}
\Orcid{???}
\Email{yanikolt@gmail.com}

\MakeArticleHeader
%========================= КОНЕЦ ШАПКИ СТАТЬИ ==============================

\pagebreak

\section{Постановка задачи}

Задачи поиска экстремумов многоэкстремальных функций часто возникают в различных областях науки, таких как химия, физика, машиностроение и т.д. В рамках данной работы решается задача минимизации многомерной функции $\varphi(y)$ в гиперинтервале \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\)  Будем предполагать, что информация о виде самой функции недоступна, целевая функция задается в формате «черного ящика». Дополнительно для целевой функции предполагаем выполнение условия Липшица, причем константа L априори неизвестна.

\begin{equation}
\label{task}
\varphi(y^*)=\min\{\varphi(y):y\in D\}
\end{equation}

\begin{equation}
\label{lip}
|\varphi(y_1)-\phi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
\end{equation}

Сведение многомерной задачи ((\ref{task})) к одномерной проводится за счет использования редукции размерности при помощи кривой Пеано [\ref{rfa:rulit:lit1}, \ref{rfa:rulit:lit2}] y(x), которая непрерывно и однозначно отображает на N-мерный гиперкуб отрезок вещественной оси [0,1]:

\begin{equation}
\label{cube}
\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
\end{equation}

\begin{equation}
\label{oneDimTask}
\varphi(y^*)=\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}
\end{equation}

 У данного способа редукции размерности имеется важное свойство, необходимое для последующих расчетов: ограниченность относительных разностей функции сохраняется, т.е. если в области $D$ функция $\varphi(y)$ удовлетворяла условию Липшица, то на интервале [0,1] функция $\varphi(y(x))$ будет удовлетворять равномерному условию Гельдера

\begin{equation}
\label{holder}
|\phi(y(x_1))-\phi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
\end{equation}

\begin{equation}
H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}
\end{equation}

 Пользуясь этим свойством, можно трактовать исходную задачу, как минимизацию одномерной гельдеровой функции $\ \ f\left(x\right)=\varphi\left(y\left(x\right)\right),\ \ x\ \in[0,\ 1]$.
Факт того, что функция задана как «черный ящик», сокращает число подходящих алгоритмов. Для исследования выбран алгоритм глобального поиска [\ref{rfa:rulit:lit3}, \ref{rfa:rulit:lit4}], который является одним из эффективных методов глобальной оптимизации, т.к. при решении многих задач опережает (по числу итераций) другие методы аналогичного назначения [\ref{rfa:rulit:lit5} – \ref{rfa:rulit:lit9}]. Алгоритм глобального поиска относится к классу характеристических алгоритмов оптимизации, хорошо масштабируется и отлично подходит для работы на параллельных вычислительных кластерах. 
В процессе своей работы алгоритм порождает последовательность точек $x^k$, в каждой из которых проходит вычисление значения минимизируемой функции $z^k=f(x^k)$. Далее под испытанием мы будем понимать подразумевать процесс вычисления значения одномерной функции, в который, помимо прочего, включается построение образа $y^k=y(x^k)$, а результатом испытания назовем пару $(x^k,\ z^k)$. Процесс распараллеливания построен таким образом, что при выполнении одной итерации метода одновременно проходят $P$ испытаний, $P\ \geq1$. Введем обозначение $k(n)$, как общее число испытаний, которые были проведены после n выполненных параллельных итераций. Приведем последовательность этапов параллельного алгоритма глобального поиска с модифицированием для решения задач с функциями, удовлетворяющими условию Гельдера:
Подготовительный этап. Изначально проводится испытание в произвольной внутренней точке $x^1\in\left[0,\ 1\right].\ $ Когда выполнено $n\geq1$ итераций, и соответствующее им число $k=k(n)$ испытаний в точках $x^i,\ 1\le\ i\le\ k$, то точки $x^{k+1},\ldots,x^{k+P}$ испытаний последующей $(n+1)$ итерации будут определять следующим образом:

 1 этап. Упорядочить по координате точки уже проведенных испытаний и граничные точки:
 \begin{equation}
\label{agp1_sort}
	0=x_0<\ x_1<\ ...\ <x_{k+1}=1 
	\end{equation}

2 этап. Вычислить текущее значение $M$, оценивающее неизвестную константу Липшица $L$:
 \begin{equation}
\label{agp2_mu}
	\mu=max\left\{\frac{|z_i-z_{i-1}|}{{{(x}_i-x_{i-1})}^{1/N}},\ i=1,\ldots,k\right\},\ M=\ \left\{\begin{matrix}r\mu,\ \mu>0,\\1,\ \mu=0,\\\end{matrix}\right.\
		\end{equation}

   где $r>1$ заданный параметр метода.

3 этап. Найти значение характеристики $R(i)$ для всех интервалов

\begin{equation}
\label{agp3_R1}
R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M}
\end{equation}

\begin{equation}
\label{agp3_Ri}
R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1
\end{equation}

   где \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).

4 этап.  Упорядочить характеристики $R\left(i\right),\ 1\ \le\ i\ \le\ k+1$ в порядке невозрастания 
\begin{equation}
\label{agp4_R_sort}
	R\left(t_1\right)\geq\ R\left(t_2\right)\geq...\geq\ R\left(t_k\right)\geq\ R(t_{k+1})\
\end{equation}

и выбрать P интервалов с номерами $t_j,\ 1\le\ j\le\ P$, значение характеристики в которых наибольшее.

5 этап. В выбранных интервалах вычислить точки $x^{k+j},\ 1\le\ j\le\ P$ и провести в них новые испытания:

\begin{equation}
\label{agp5_x1}
	x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\ t_j=1,\ t_j=k+1
\end{equation}

\begin{equation}
\label{agp4_xi}	
	x^{k+1}=\frac{x_{t_j}+x_{t_j-1}}{2}-sign\left(z_{t_j}-z_{t_j-1}\right)\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N,\ 1<t_j<k+1
\end{equation}

Важно отметить, что в рассматриваемых задачах самый трудоемкий и времязатратный этап алгоритма – это вычисление значения функции в точке. Поэтому параллельное вычисление этих значений положительно влияет на скорость работы алгоритма.

Используемые критерии остановки:

	по длине интервала, как только происходит выполнение условия \(\Delta_{t_j}\leqslant \varepsilon\) хотя бы для какого-то номера $t_j,\ 1\le\ j\le\ P$. Основной критерий остановки, используется в прикладных задачах оптимизации и в тестовых наборах, для которых точка глобального минимума заранее неизвестна.

	по попаданию в окрестность глобального минимума, $\left|x_{t_j}-\ x^\ast\right|<\varepsilon,\ \ t_j,\ 1\le\ j\le\ P$, где $x^\ast$ -- точка глобального минимума, если ее значение известно. Используется в тестовых задачах, когда $x^\ast$ известна до начала работы алгоритма.

Как оценка глобально-оптимального решения рассматриваемой задачи (1) выбираются значения

\begin{equation}
f_k^*=\min_{1\leqslant i \leqslant k}f(x_i), x_k^*=arg \min_{1\leqslant i \leqslant k}f(x_i)
\end{equation}

\section{Реализация с использованием Intel oneAPI}

Рассматриваемые многомерные многоэкстремальные задачи обладают высокой трудоемкостью численного решения, поскольку при увеличении размерности задачи наблюдается экспоненциальный рост вычислительных затрат. Вместе с тем быстрый процесс развития современных вычислительных средств, включая распараллеливание, предоставляет всё больше различных новых возможностей для решения проблем оптимизации. Однако на фоне этого возникает задача эффективного распараллеливания, а многообразие различных типов современных ускорителей и средств разработки для них дают пользователям немалый выбор [\ref{rfa:rulit:lit10}].

Но перейдя от частной задачи к общей ситуации, разберем, какие предлагаются варианты решения. Для обеспечения высокой производительности вычислений в рамках каких-то новых рабочих задач требуются различные вычислительные архитектуры. Например, компанией Intel предлагается следующая классификация ускорителей: скалярные (CPU), векторные (GPU), матричные и ПЛИС (FPGA). Эти классы отличаются архитектурой, которая прямым образом влияет на их функционал.

Рассмотрим их по порядку. Скалярные (CPU), являются кэш-ориентированными, основаны на оптимизации однопоточной производительности и на решение задач общего назначения. Несмотря на то, что большинство процессоров сейчас многоядерные, достаточно много ресурсов тратится на обеспечение их однопоточной производительности. В GPU большая часть ресурсов задействована под вычисления [\ref{rfa:rulit:lit11}], а не под кэш, как в случае с CPU, поэтому в GPU применима стратегия SIMD[\ref{rfa:rulit:lit12}].   Третий класс -- это матричные процессоры, рассчитанные на быструю работу с матрицами. Прежде всего это ускорители из области искусственного интеллекта, нейронные процессоры, например, процессоры машинного зрения, тензорные и другие. И последний класс -- программируемые логические интегральные схемы (ПЛИС), FPGA [\ref{rfa:rulit:lit13}]. Основу структуры составляет матрица логических элементов, функции этих элементов и связи между ними могут модифицироваться – программироваться, в процессе использования. Сфера применения ПЛИС достаточно широка, они используются в бытовой электронике, телекоммуникационном оборудовании, разнообразной робототехнике и при прототипировании микросхем.

Однако использование преимуществ нескольких типов архитектур является сложной задачей для разработчиков. Для каждой архитектуры требуются разные языки, отдельные инструменты, а повторное использование кода ограничено. Это делает разработку сложной, дорогостоящей и трудоемкой.

Например, когда возникает потребности выполнения алгоритма на нескольких типах ускорителях, то при написании кода программы, возникают проблемы связанные с несовместимости разных средств разработки и разными архитектурными особенностиями. Для более четкого понимания этого, проведем небольшой обзор инструментов и средств разработки для программирования ускорителей. Для GPU часто используются графические API и шейдерные языки: DirectX, OpenGL, Vulcan, Metal. Некоторые производители, к примеру, NVidia и AMD, создают свои специальные средства, которые подходят под их ускорители (NVidia CUDA, AMD ROCm). Для программирования под FPGA применяются языки описания архитектур, например, Verilog и VHDL. Также есть набор общих средств, которые отличаются по применимости, сложности написания кода, текущей поддержке: OpenMP, OpenCL, Python, SYCL и другие.

Как правило, если остановиться на одной из вычислительных архитектур, то для возможности запуска на абсолютно другой может потребоваться адаптация части кода, а может даже придется написать программу с нуля, применяя другие инструменты. В большинстве случаев, будет создан новый проект, и необходимо будет поддерживать несколько программ, в которых используются разные технологии.

Для создания универсального кода работающего на различных устройствах можно воспользоваться набором инструментов Intel oneAPI. Он прост, открыт и позволяет разработчику обеспечивать высокую производительность в разных архитектурах. А поскольку oneAPI основан на стандартах и открытых спецификациях, риски при переносе снижаются. Это дает возможность один раз написать код и в дальнейшем запустить его на другом устройстве. Также к преимуществам oneAPI можно отнести возможность применения в различных прикладных решениях, например, в задачах машинного обучения.

В рамках данной задачи необходимы возможности распараллеливания, которые обеспечиваются за счет включения в oneAPI языка Data Parallel C++ и набора библиотек, облегчающих межархитектурную разработку. Data Parallel C++ основан на широко известном языке С++ и включает в себя SYCL от группы Khronos и расширения от комьюнити.

Все это позволяет повторно использовать код в разных архитектурах и выполнять пользовательскую настройку ускорителей. Что дает разработчикам гибкость, позволяющую отказаться от патентованных подходов, и открывает возможности для использования аппаратных средств, которые ранее были невозможны.

Руководствуясь приведенным ранее алгоритмом, реализована возможность одновременно вычислять сразу несколько значений целевой функции, используя при этом инструменты распараллеливания Intel oneAPI. Также имеется возможность выбрать устройство, на котором будут проходить вычисления значений функции. Остальные этапы рассмотренного алгоритма необходимо проводить последовательно, потому что в процессе их выполнения проходит работа с достаточно большим количеством накопленной ранее поисковой информации, в связи с этим реализуем их на CPU.

Таким образом, первые этапы рассматриваемого алгоритма выполняются на центральном процессоре (CPU). Далее, полученные в ходе выполнения одной итерации новые P координат из интервалов с наибольшими характеристиками, передаются с помощью промежуточного буфера на устройство (CPU, GPU, FPGA), выбранное для исполнения распараллеленного блока кода, для вычисления значения функции в них. Затем найденные значения функции в этих точках передаются через промежуточный буфер обратно на центральный процессор. Общая схема организации параллельных вычислений приведена на рис. \ref{fig:s1}.

\begin{figure}
\begin{center}
  \includegraphics[width=0.7\linewidth]{./pic/s1.pdf}
  \caption{Общая схема организации параллельных вычислений.}
  \label{fig:s1}  
\end{center}
\end{figure}

\section{Результаты численных экспериментов}

В рамках работы были проведены вычислительные эксперименты на персональном компьютере с процессором Intel Core i5-10600 3.3GHz, 32 GB оперативной памяти и встроенной графической картой Intel UHD Graphics 630. Для получения исполняемого программного кода использовался компилятор Intel oneAPI DPC++ 2021.1. Вычислительные эксперименты выполнялись с использованием разработанного программного комплекса Globalizer [\ref{rfa:rulit:lit14}, \ref{rfa:rulit:lit15}].

Большинство известных тестовых задач из области многомерной глобальной оптимизации характеризуются небольшим временем вычисления значений целевой функции. Из-за этого при проведении экспериментов бывает сложно понять с чем может быть связано относительно большое время выполнения параллельного алгоритма: с появившимися, сравнительно с последовательной версией, накладными расходами или причина в неэффективном распараллеливании. В реальных же задачах вычисление функции самая трудоемкая и время затратная операция, поэтому такой проблемы не возникает.

Нами предлагается вычислительно трудоемкая модификация существующих тестовых задач, расчет целевой функции заключается в интегрировании исходной тестовой функции по части параметров. Для этого изначально порождается задача, размерность которой в два раза больше искомой. При вычислении значения функции, первые $N$ координат фиксируются, а по остальным производится численное интегрирование по области определения функции. Для интегрирования используется метод средних прямоугольников.

\begin{equation}
\varphi \left( {{y}_{1}},\ldots ,{{y}_{N}}~ \right)=\underset{{{a}_{N+1}}}{\overset{{{b}_{N+1}}}{\mathop \int }}\,\underset{{{a}_{N+2}}}{\overset{{{b}_{N+2}}}{\mathop \int }}\,\ldots \underset{{{a}_{2N}}}{\overset{{{b}_{2N}}}{\mathop \int }}\,\psi \left( {{y}_{1}},\ldots ,{{y}_{N}},{{y}_{N+1}},~..,{{y}_{2N}}~ \right)~d{{y}_{N+1}}d{{y}_{N+2}}..d{{y}_{2N}}~= 
\end{equation}
\begin{displaymath}
\underset{{{i}_{1}}=0}{\overset{M-1}{\mathop \sum }}\,\underset{{{i}_{2=0}}}{\overset{M-1}{\mathop \sum }}\,..\underset{{{i}_{N}}=0}{\overset{M-1}{\mathop \sum }}\,\left( \left( \underset{j=1}{\overset{2N}{\mathop \prod }}\,{{h}_{j}} \right)\psi \left( {{y}_{1}},{{y}_{2}},~..,{{y}_{N}},~{{y}_{{{\left( N+1 \right)}_{{{i}_{1}}}}}},{{y}_{{{\left( N+2 \right)}_{{{i}_{2}}}}}}~,~..,{{y}_{{{\left( 2N \right)}_{{{i}_{N}}}}}} \right) \right)
\end{displaymath}
\begin{displaymath}
D=\left\{ y~\in ~{{R}^{2N}}:~{{a}_{i}}\le ~~{{y}_{i}}\le ~~{{b}_{i}},~1\le i\le 2N \right\}
\end{displaymath}
\begin{equation}
{{y}_{{{\left( N+k \right)}_{{{i}_{k}}}}}}={{a}_{k}}+{{i}_{k}}*{{h}_{k}}+\frac{{{h}_{k}}}{2}
\end{equation}
\begin{equation}
{{h}_{k}}=~\frac{{{b}_{k}}-~{{a}_{k}}}{M},
\end{equation}

где $M$ -- количество участков интегрирования по одной координате, а $\psi$ – исходная тестовая функция. Очевидно, чем больше значение $M$, тем больше проводится вычислений. Изменяя число узлов в сетке интегрирования по всем координатам или число участков по одной, можно регулировать время выполнения вычислений.

Вначале проведем вычислительные эксперименты на классической задаче – функции Растригина. Она задается простой формулой, сепарабельна и не имеет ограничений по размерности.

\begin{equation}
\label{Rastrigin}
\psi \left( {{y}_{1}},\ldots ,{{y}_{2N}}~ \right)=~\underset{i=1}{\overset{2N}{\mathop \sum }}\,(y_{i}^{2}-10\cos \left( 2\pi {{y}_{i}} \right)+10)
\end{equation}

\begin{displaymath}
-2.2<y\_i<1.8~,~1\le i\le 2N
\end{displaymath}

На рис. \ref{fig:s2} изображены линии уровня двухмерной функции Растригина (слева) и интегрированной четырехмерной функции Растригина (справа). Как можно видеть, новая функция не сильно отличается от двухмерного оригинала, и даже глобальный минимум остался прежним – точка (0, 0).

\begin{figure}
\begin{center}
  \includegraphics[width=1.0\linewidth]{./pic/s2.png}
  \caption{Линии уровня двумерной функции Растригина (слева) и интегрированной четырехмерной функции Растригина (справа).}
  \label{fig:s2}  
\end{center}
\end{figure}

Число итераций параллельного алгоритма глобального поиска (ПАГП) было ограничено 1000000, точность поиска $\varepsilon=0.01$ и параметр метода $r\ =\ 3$. Размерность задачи $N = 4$ и $5$. При вычислениях на CPU число потоков варьировалось от 1 до 8, а при вычислении на GPU от 256 до 1024. Использовался критерий остановки по попаданию в окрестность. В таблице \ref{table1} приведено число итераций ПАГП, в таблице \ref{table2} – ускорение по сравнению с однопоточным запуском.

\begin{table}[!ht]
    \centering
            \caption{Число итераций ПАГП при решении интегрированной функции Растригина.}
    \label{table1}
    \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
    \hline
        N & CPU &   & ~ & ~ & ~ & GPU  & ~ & ~ \\ \hline
        ~ & P=1 & P=2 & P=4 & P=8 &   & P=256 & P=512 & P=1024  \\ \hline
        4 & 61231 & 26592 & 21007 & 6728 &   & 340 & 301 & 92  \\ \hline
        5 & 703548 & 328141 & 258351 & 99524 &   & 4642 & 2194 & 995  \\ \hline
    \end{tabular}
\end{table}

\begin{table}[!ht]
    \centering
            \caption{Ускорение ПАГП при решении интегрированной функции Растригина}
    \label{table2}
    \begin{tabular}{|l|l|l|l|l|l|l|l|}
    \hline
        N & CPU &   & ~ & ~ & GPU  & ~ & ~ \\ \hline
        ~ & P=2 & P=4 & P=8 & ~ & P=256 & P=512 & P=1024  \\ \hline
        4 & 2.1 & 2.1 & 5.6 & ~ & 2.8 & 3.1 & 9.4  \\ \hline
        5 & 1.9 & 2.0 & 4.3 & ~ & 2.4 & 4.9 & 9.7  \\ \hline
    \end{tabular}
\end{table}

Как можно видеть из полученных результатов, использование инструментов Intel oneAPI, при распараллеливании алгоритма глобального поиска, показала хорошие результаты на тестовой функции.

Далее проведем эксперименты на серии задач полученных интегрированием функций из генератора GKLS. Данный генератор описан в работах [\ref{rfa:rulit:lit16}, \ref{rfa:rulit:lit17}], дает возможность создавать серии задач многоэкстремальной оптимизации и заранее задавать их свойства, такие как: размерность задачи, количество локальных минимумов, размеры их областей притяжения, координата точки глобального минимума, значение функции в ней и т.п. На рис. \ref{fig:s3} изображены линии уровня двухмерной функции GKLS номер 9 (слева) и интегрированной функции GKLS номер 9, полученной интегрированием четырехмерной функции по последним двум параметрам (справа).

\begin{figure}
\begin{center}
  \includegraphics[width=1.0\linewidth]{./pic/s3.png}
  \caption{Функции GKLS (слева) и интегрированной функции GKLS (справа).}
  \label{fig:s3}  
\end{center}
\end{figure}

Число итераций ПАГП было ограничено 100000, точность поиска $\varepsilon=0.01$ и параметр метода $r\ =\ 4$. Размерность задачи $N = 4$ и 5. При вычислениях на CPU число потоков варьировалось от 1 до 8, а при вычислении на GPU от 256 до 1024. В таблице \ref{table3} приведено ускорение по сравнению с однопоточным запуском.

\begin{table}[!ht]
    \centering
    \caption{Ускорение ПАГП при решении серии интегрированных функций GKLS.}
    \label{table3}
    \begin{tabular}{|l|l|l|l|l|l|l|l|}
    \hline
        N & CPU &   & ~ & ~ & GPU  & ~ & ~ \\ \hline
        ~ & P=2 & P=4 & P=8 & ~ & P=256 & P=512 & P=1024  \\ \hline
        4 & 1.9 & 3.0 & 1.5 & ~ & 1.2 & 2.3 & 4.5  \\ \hline
        5 & 1.9 & 3.0 & 1.6 & ~ & 1.2 & 2.3 & 4.5  \\ \hline
    \end{tabular}
\end{table}

%=======================  Список литературы ==================================
%
% Команда \LITERRUS печатает "СПИСОК ЛИТЕРАТУРЫ" и обнуляет счетчики
%
\LITERRUS
% \rlitem{arg1}{arg2} --- создает пункт списка литературы.
% arg1 --- символическая ссылка на источник, используется при создании ссылки командой \ref
% arg2 --- текст, помещаемый в пункт списка. Может содержать команды выбора начертания 
%          символов и др.

\rlitem{rfa:rulit:lit1}{
\textit{Sergeyev, Ya.D.} Introduction to global optimization exploiting space-filling curves / Ya.D. Sergeyev, R.G. Strongin, D. Lera – Springer, 2013. 125 p.}

\rlitem{rfa:rulit:lit2}{
\textit{Strongin, R.G.} Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms / R.G. Strongin, Ya.D. Sergeyev – Kluwer Academic Publishers, 2000. 704 p.}

\rlitem{rfa:rulit:lit3}{
\textit{Стронгин, Р.Г.} Параллельные вычисления в задачах глобальной оптимизации / Р.Г. Стронгин, В.П. Гергель, В.А. Гришагин, К.А. Баркалов – М.: Издательство Московского университета, 2013. 280 с.}

\rlitem{rfa:rulit:lit4}{
\textit{Pinter (Ed.), J.D.} Global Optimization: Scientific and Engineering Case Studies / J.D. Pinter – Springer, 2006. 546 p.}

\rlitem{rfa:rulit:lit5}{
\textit{Стронгин, Р.Г.} Параллельные методы решения задач глобальной оптимизации / Р.Г. Стронгин, В.П. Гергель, К.А Баркалов, // Известия высших учебных заведений. Приборостроение, 2009. Т. 52. № 10. С. 25 – 33.}

\rlitem{rfa:rulit:lit6}{
\textit{Захарова, Е.} Обзор методов многомерной оптимизации. / Е.М. Захарова, И.К. Минашина, // Информационные процессы, 2014. Т 14. №3. С. 256 – 274.}

\rlitem{rfa:rulit:lit7}{
\textit{Gaviano, M.} Software for generation of classes of test functions with known local and global minima for global optimization/ M. Gaviano, D. Lera, D. E. Kvasov, Y. D. Sergeyev // ACM Transactions on Mathematical Software. – 2003. – Vol.  29. – P. 469-480.}

\rlitem{rfa:rulit:lit8}{
\textit{Сергеев, Я.Д.} Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. – 352 c.}

\rlitem{rfa:rulit:lit9}{
\textit{Gablonsky, J.M.} A Locally-Biased Form of the DIRECT Algorithm / J.M.  Gablonsky, C.T. Kelley, // Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.}

\rlitem{rfa:rulit:lit10}{
\textit{Paulavicius, R.} Parallel branch and bound for global optimization with combination of Lipschitz bounds / R. Paulavicius, J. Zilinskas and A. Grothey // Optimization Methods \& Software,  2011. – Vol. 26, No. 3. P. 487–498.}

\rlitem{rfa:rulit:lit11}{
\textit{Боресков А.А.} Параллельные вычисления на GPU. Архитектура и программная модель CUDA / Боресков А.А., Харламов А.А., Марковский Н.Д., Микушин Д.Н., Мортиков Е.В., Мыльцев А.А., Сахарных Н.А., Фролов В.А. – М.:  Издательство Московского университета, 2015. 336 с.}

\rlitem{rfa:rulit:lit12}{
\textit{Таненбаум, Э.} Архитектура компьютера / Э. Таненбаум, Т. Остин; перевод с английского Е. Матвеева. — 6-е изд. — Санкт-Петербург [и др.]: Питер, 2014. 816 с.}

\rlitem{rfa:rulit:lit13}{
\textit{Комолов Д.А.} Системы автоматизированного проектирования фирмы Altera MAX+Plus II и Quartus II. / Комолов Д.А. Мяльк Р.А. Зобенко А.А. Филиппов А.С., Издательство РадиоСофт, 2002. 355 c.}

\rlitem{rfa:rulit:lit14}{
\textit{Gergel, V.P.} A novel supercomputer software system for solving time-consuming global optimization problems / V.P. Gergel, K.A. Barkalov A.V. Sysoyev // Numerical Algebra, Control and Optimization, 2018. Vol. 8, No. 1. P. 47-62.}

\rlitem{rfa:rulit:lit15}{
\textit{Sysoyev, A., Barkalov, K., Sovrasov, V., Lebedev, I., Gergel, V.} AGS NLP solver. \url{ https://github.com/sovrasov/ags_nlp_solver}}

\rlitem{rfa:rulit:lit16}{
\textit{Сергеев, Я.Д.} Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. 352 c.}

\rlitem{rfa:rulit:lit17}{
\textit{Gaviano, M.} Software for generation of classes of test functions with known local and global minima for global optimization/ M. Gaviano, D. Lera, D. E. Kvasov, Y. D. Sergeyev // ACM Transactions on Mathematical Software, 2003. Vol. 29. P. 469-480.}

%=======================  Список литературы ==================================
\medskip
\DateRU

\medskip

\MakeAuthorsInfoRU

%===========================  REFERENCES =====================================
\pagebreak

\REFERENCES

\elitem{rfa:enlit:COPE1}{
Guidelines on Good Publication Practice. Committee on Publication Ethics (COPE).
\url{https://publicationethics.org/files/u7141/1999pdf13.pdf}.
Cited November 12, 2021.}

\elitem{rfa:enlit:Kotelnikov}{
I.~A.~Kotel'nikov and P.~Z.~Chebotaev, 
\textit{\LaTeX\ in Russian} 
(Sibirskiy Khronograph, Novosibirsk, 2004) [in Russian].}

\elitem{rfa:enlit:Samarin}{
M.~Goossens, F.~Mittelbach, and A.~Samarin, 
\textit{The \LaTeX~Companion.} 
(Addision--Wesley Professional, 1994; Mir, Moscow, 1999) 
[in Russian].}

\elitem{rfa:enlit:Lvovski}{
S.~M.~L'vovsky. \textit{Typesetting and Layout in the \LaTeX~System}
(MCNMO Publ., Moscow, 2003) 
[in Russian].}

\elitem{rfa:enlit:rot}{
Ya.~S.~Bugrov and S.~M.~Nikol'sky, 
\textit{Higher Mathematics: Differential Equations. Multiple Integrals. Ranks} 
(Phoenix Publ., Rostov-na-Donu, 1998) 
[in Russian].}

\elitem{rfa:enlit:Antonov}{
A.~S.~Antonov, 
\textit{Parallel Programming Using OpenMP Technology}
(Mosk. Gos. Univ., Moscow, 2009) 
[in Russian].}

\elitem{rfa:enlit:Lapko}{
O.~Lapko, 
\textit{The Makecell Package}. 
\url{https://mirror.truenetwork.ru/CTAN/macros/latex/contrib/makecell/makecell-rus.pdf}.
Cited November 12, 2021 [in Russian].}

\elitem{rfa:enlit:mrow}{%
Multirow~--- Create Tabular Cells Spanning Multiple Rows.
\url{https://www.ctan.org/pkg/multirow}.
Cited November~12, 2021.}

\medskip
\DateEN

\bigskip

\MakeAuthorsInfoEN

%\end{document}
%============================ КОНЕЦ СТАТЬИ ==================================

\end{document}

### 2022-09-04 2022/OPTIMA/paper.tex
all the reviewers' comments are fixed

@@
 %
-\title{An accelerated algorithm for finding efficient solutions in multiobjective problems with black-box multiextremal criteria\thanks{The work was supported by the Ministry of Science and Higher Education of the Russian Federation (project no.0729-2020-0055) and by the Research and Education Mathematical Center (project no.
+\title{An accelerated algorithm for finding efficient solutions in multiobjective problems with black-box multiextremal criteria\thanks{The work was supported by the Ministry of Science and Higher Education of the Russian Federation (project No.
+0729-2020-0055) and by the Research and Education Mathematical Center (project No.
 075-02-2020-1483/1).}}
@@
 %However, the function $f(x(t))$ will not meet the Lipschitz if $f(x)$ is Lipschitzian. It will satisfy the H{\"o}lder condition
-However, in the case when the function $f(x)$ is Lipschitzian this property is not valid for the reduced function. The function $f(x(t))$ will satisfy the H{\"o}lder condition
+However, in the case when the function $f(x)$ is Lipschitzian the function $f(x(t))$ will satisfy the H{\"o}lder condition
 \begin{equation}\label{holder}

### 2022-09-04 2022/PAVT Special Issue/template.tex

@@
 
-In the complex problems of chemical kinetics, kinetic constants of reactions cannot be found analytically, they can only be found using numerical methods (see, for example, [references]).
+In the complex problems of chemical kinetics, kinetic constants of reactions cannot be found analytically, they can only be found using numerical methods (see, for example, \cite{Zaynullin2020,Enikeeva2020,Koledina2019}).
 In this case, the quality criterion (objective function) is set in the form of a certain algorithm, which includes the stage of numerical modeling, which, in turn, makes it compute-intensive.
@@
 
-This paper reflects the results of applying parallel methods of Lipschitz optimization developed in [refernces], to the solution of inverse problems of chemical kinetics.  
+This paper reflects the results of applying parallel methods of Lipschitz optimization developed in \cite{Barkalov2016,Strongin2018,Sysoyev2017}, to the solution of inverse problems of chemical kinetics.
 The main part of the paper is structured as follows. 

### 2022-09-22 2022/PAVT Special Issue/template.tex
submitted version of the paper

@@
 % Full title of the paper (Capitalized)
-\Title{Using the ... Algorithm to Detect Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction}
+\Title{On Solving the Problem of Finding Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction Using Parallel Global Search Algorithm}
 % Дежурное название: Использование ...алгоритма для определения кинетических параметров процесса каталитической изомеризации пентан-гексановой фракции
 % Еще один вариант: О решении задачи определения Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction с помощью Parallel Global Search Algorithm
-% На английском: About Solving the Problem of Finding Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction Using Parallel Global Search Algorithm
+% На английском: On Solving the Problem of Finding Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction Using Parallel Global Search Algorithm
 
 % MDPI internal command: Title for citation in the left column
-\TitleCitation{Title}
+\TitleCitation{On Solving the Problem of Finding Kinetic Parameters}
 
@@
 \newcommand{\orcidauthorA}{0000-0001-5273-2471} % Add \orcidA{} behind the author's name
-\newcommand{\orcidauthorB}{0000-0002-8736-0652} % Add \orcidB{} behind the author's name
-\newcommand{\orcidauthorC}{0000-0003-4219-4870}
+\newcommand{\orcidauthorB}{0000-0001-6776-0096}
+\newcommand{\orcidauthorC}{0000-0002-8736-0652} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorD}{0000-0001-6073-9585} 
+\newcommand{\orcidauthorE}{0000-0003-4219-4870}
 
 % Authors, for the paper (add full first names)
-\Author{Konstantin Barkalov $^{1}$\orcidA{}, Irek Gubaydullin $^{2,3}$, Ilya Lebedev $^{1}$\orcidB{}, Roza Faskhutdinova $^{2,3}$ and Leniza Enikeeva $^{3,4,}$*\orcidC{}}
+\Author{Konstantin Barkalov $^{1}$\orcidA{}, Irek Gubaydullin $^{2,3}$, Evgeny Kozinov$^{1}$\orcidB{}, Ilya Lebedev $^{1}$\orcidC{}, Roza Faskhutdinova\orcidD{} $^{2,3}$, Azamat Faskhutdinov$^{2}$ and Leniza Enikeeva $^{3,4,}$*\orcidE{}}
 
@@
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Konstantin Barkalov, Irek Gubaydullin, Ilya Lebedev, Roza Faskhutdinova and Leniza Enikeeva}
+\AuthorNames{Konstantin Barkalov, Irek Gubaydullin, Evgeny Kozinov, Ilya Lebedev, Roza Faskhutdinova, Azamat Faskhutdinov, Leniza Enikeeva}
 
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Barkalov, B.; Gubaydullin, I.; Lebedev, I.; Faskhutdinova, R.; Enikeeva, L.}
+\AuthorCitation{Barkalov, B.; Gubaydullin, I.; Kozinov, E.; Lebedev, I.; Faskhutdinova, R.; Faskhutdinov, A.; Enikeeva, L.}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{A single paragraph of about 200 words maximum.
-For research articles, abstracts should give a pertinent overview of the work.
-We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusions: indicate the main conclusions or interpretations.
-The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}
+\abstract{The article is devoted to the problem of developing a kinetic model of a complex chemical reaction using a parallel optimization method.
+The design of the kinetic model consists in finding the kinetic parameters of the reaction, which cannot be calculated analytically, and since the chemical reaction involves many stages, the optimization problem is multi-extreme.
+As a chemical reaction, the process of catalytic isomerization of the pentane-hexane fraction is considered, which is now important due to the switch of the oil refining industry to the production of gasoline corresponding to the Euro-5 standard.
+To solve the optimization problem, the authors have developed parallel global search algorithm and a program based on Lipschitz optimization.
+The kinetic parameters found made it possible to develop a mathematical model of the process, which gives good agreement with industrial data.}
 
 % Keywords
-\keyword{kinetic model; inverse problems of chemical kinetics; mathematical modeling; catalytic isomerisation of pantane-hexane fraction (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}
+\keyword{kinetic model; inverse problems of chemical kinetics; mathematical modeling; catalytic isomerisation of pentane-hexane fraction; Lipschitz optimization; black-box functions; parallel computing}
 
@@
 
-%Часть ННГУ - начало
+Due to the transition of the domestic oil refining industry to the production of gasoline conforming to the Euro-5 standard, an urgent task is to reduce the content of aromatic hydrocarbons and, in particular, benzene in motor fuel while maintaining the octane value.
+In connection with the transition of the oil refining industry to the production of motor gasoline conforming to Euro-5 and Euro-6 standards, the reduction of aromatic hydrocarbons in motor fuel and, mainly, benzene becomes an urgent task.
+At the same time, removing aromatic hydrocarbons, which are high-octane components, it is necessary to preserve the octane number of gasoline.
+It is known that hydrocarbons of branched structure have a higher octane number than hydrocarbons of linear structure.
+For example, in $n$-pentane, the octane number according to the research method is 62 points, and in $iso$-pentane it is 93 points.
+On an industrial scale, catalytic isomerization units of light paraffins have been introduced to produce hydrocarbons of an isomeric structure.
+The catalytic isomerization of light paraffins, in turn, makes it possible to obtain a high-octane component of automobile gasoline with a minimum content of aromatic hydrocarbons.
+The main advantage of this process is the production of high-octane gasoline, while aromatic hydrocarbons are practically not formed.
+Straight-run gasoline fractions with boiling limits of 62-70 $^0$C containing mainly $n$-pentane and $n$-hexane, pentane-hexane fractions from gas condensate or from gas fractionation plants, as well as catalytic reforming raffinates are used as feedstock for catalytic isomerization.
+Modern bifunctional catalysts are used to carry out the process.
+In order to prevent coke deposition, the catalytic isomerization process is carried out in a hydrogen-containing gas environment under pressure.
+
+Computer information technologies are widely used in the modeling of production processes and today it is difficult to imagine research activities without the use of modern means of constructing and using models.
+Moreover, special attention is currently being paid to the development of a detailed kinetic model of complex industrial processes.
+Based on these models, it is possible to significantly increase the efficiency of the process, while minimally spending energy and material resources.
+The main advantage of such methods is the absence of additional costs for the construction of pilot plants.
+The developed model should describe the process within a wide range of operating conditions, have high flexibility, allowing for changes, calculations of the composition of the reaction mixture, etc.
+
+In the article \cite{enik2021} we have already reviewed this process, but in order to simplify the model, the scheme of chemical transformations, in particular hydrocracking reactions, have been revised.
+Thus, the most probable hydrocracking reactions based on the carbonium-ion mechanism were formed.
+The new transformation scheme implies the recalculation of the kinetic parameters of the reaction, which form a mathematical model of the process.
+
+In the article \cite{enik2021}, a genetic algorithm was used to calculate kinetic parameters, and the calculation time of kinetic parameters was about 10 hours.
+Therefore, in order to develop a mathematical model of a new transformation scheme, the question of choosing a more effective optimization method arises.
 
@@
 
-Numerical methods for solving such multiextremal problems (global optimization methods) differ significantly from local optimization methods since they require the study of the entire search domain (see, for example, \cite{PaulaviciusZilinskas2014,Sergeyev2017}).
-These methods can be divided into two classes: meta-heuristic and deterministic.
+Computational algorithms for solving such multiextremal problems (global optimization methods) differ vastly from local optimization techniques since they require the study of the entire search domain (see, for example, \cite{PaulaviciusZilinskas2014,Sergeyev2017}).
+These methods can be divided into two classes: metaheuristic and deterministic.
 Metaheuristic algorithms are usually based on simulations of processes occurring in the wild.
-Examples of metaheuristic algorithms are simulated annealing, evolution and genetic algorithms, etc. (see, for example, \cite{Battiti2009,Eiben2015}).
+Examples of metaheuristic algorithms are simulated annealing, evolutionary computing, etc. (see, for example, \cite{Battiti2009,Eiben2015}).
 These algorithms are included in many software libraries for solving global optimization problems (for example, MATLAB Global Optimization Toolbox, Scipy.Optimize, etc.).
@@
 The main part of the paper is structured as follows.
-Section 2 provides a description of the mathematical model of the chemical reaction under study.
-Section 3 discusses the standard formulation of the Lipschitz global optimization problem and the algorithm for solving it.
-It also outlines the issues related to the technical aspects of the implementation of the parallel algorithm.
-The results of the numerical solution of the inverse problem of chemical kinetics using the developed parallel algorithm are discussed in Section 4.
-The last section contains the conclusion and plans for future work.
 
-%Часть ННГУ - конец
+Section \ref{sec_matmodel} provides a description of the mathematical model of the chemical reaction under study.
+Section \ref{sec_GSA} discusses the standard formulation of the Lipschitz global optimization problem and the algorithm for solving it.
+Section \ref{sec_numexper} outlines the issues related to the technical aspects of the implementation of the parallel algorithm.
+The physico-chemical interpretation of the results of numerical calculations using the developed parallel algorithm is presented in the Section \ref{sec_discussion}.
+%The last section contains the conclusion and plans for future work.
 
-\section{Mathematical model}
+\section{Mathematical model}\label{sec_matmodel}
 
-Let's consider the scheme of chemical transformations of the isomerization of the pentane-hexane fraction. The reaction consists of 48 reaction stages, ...
-
-2-MP is 2-methyl pentane;
-
-3-MP is 3-methyl pentane;
-
-2,2-DMB is 2,2-dimethyl butane;
-
-2,3-DMB is 2,3-dimethyl butane
-
-CH* is cyclohexane
-
-MCP is methyl cyclopentane
-
-B is benzene.
-
-3-MP is 3-methyl pentane;
-
-2,2-DMB is 2,2-dimethyl butane;
-
-2,3-DMB is 2,3-dimethyl butane
-
-CH* is cyclohexane
+Let's consider the mechanism of chemical transformations of the isomerization of the pentane-hexane fraction.
+The reaction consists of 48 reaction stages, which are presented in the table \ref{tab_chemistry}.
+The following abbreviations were adopted: 2-MP is 2-methylpentane; 3-MP is 3-methylpentane; 2,2-DMB is 2,2-dimethylbutane; 2,3-DMB is 2,3-dimethylbutane; CH* is cyclohexane; MCP is methyl cyclopentane; B is benzene; CP is cyclopentane.
 
 \begin{table}[H] 
-\caption{Transformations of the catalytic isomerization of the pentane-hexane fraction.\label{tab11}}
+\caption{Chemical transformations of the catalytic isomerization of the pentane-hexane fraction.\label{tab_chemistry}}
 \newcolumntype{C}{>{\centering\arraybackslash}X}
-\begin{tabularx}{\textwidth}{CCCC}
+\begin{tabularx}{\textwidth}{CC}
 \toprule
-\textbf{No}	& \textbf{Reaction}	& \textbf{E, kJ/mol}     & \textbf{log($k_0$)} \\
+ \textbf{Reaction stages} & \textbf{Reaction stages} \\
 \midrule
-1	& $n$-C$_5$H$_{12}$ $\rightarrow$ $iso$-C$_5$H$_{12}$ & 149.320 & 18.492\\
-2 & $iso$-C$_5$H$_{12}$ $\rightarrow$ $n$-C$_5$H$_{12}$ & 160.474 & 19.507 \\
-3 & $n$-C$_6$H$_{14}$ $\rightarrow$ 2-MP & 102.833 & 8.127 \\
-4 & 2-MP $\rightarrow$ $n$-C$_6$H$_{14}$ & 106.317 & 11.279 \\
-
+1) $n$-C$_5$H$_{12}$ $\rightarrow$ $iso$-C$_5$H$_{12}$ & 25) $n$-C$_6$H$_{14}$ $\rightarrow$ CH* + H$_2$\\
+2) $iso$-C$_5$H$_{12}$ $\rightarrow$ $n$-C$_5$H$_{12}$ & 26) CH* + H$_2$ $\rightarrow$ $n$-C$_6$H$_{14}$ \\
+3) $n$-C$_6$H$_{14}$ $\rightarrow$ 2-MP & 27) 2-MP $\rightarrow$ MCP+ H$_2$ \\
+4) 2-MP $\rightarrow$ $n$-C$_6$H$_{14}$ & 28) MCP + H$_2$ $\rightarrow$ 2-MP  \\
+5) $n$-C$_6$H$_{14}$ $\rightarrow$ 3-MP & 29) 3-MP $\rightarrow$ MCP + H$_2$ \\
+6) 3-MP $\rightarrow$ $n$-C$_6$H$_{14}$  & 30) MCP + H$_2$ $\rightarrow$ 3-MP \\
+7)  2-MP $\rightarrow$ 3-MP & 31) 2,2-DMB $\rightarrow$ MCP + H$_2$ \\
+8) 3-MP $\rightarrow$ 2-MP  & 32) MCP + H$_2$ $\rightarrow$ 2,2-DMB \\
+9) 2-MP $\rightarrow$ 2,2-DMB  & 33) 2,3-DMB $\rightarrow$ MCP + H$_2$ \\
+10) 2,2-DMB $\rightarrow$ 2-MP  & 34) MCP + H$_2$ $\rightarrow$ 2,3-DMB \\
+11) 2-MP $\rightarrow$ 2,3-DMB  & 35) CP + H$_2$ $\rightarrow$ $n$-C$_5$H$_{12}$ \\
+12) 2,3-DMB $\rightarrow$ 2-MP  & 36) $n$-C$_5$H$_{12}$ $\rightarrow$ CP + H$_2$ \\
+13) 3-MP $\rightarrow$ 2,2-DMB  & 37) $n$-C$_4$H$_{10}$ $\rightarrow$ $iso$-C$_4$H$_{10}$ \\
+14) 2,2-DMB $\rightarrow$ 3-MP  & 38) $iso$-C$_4$H$_{10}$ $\rightarrow$ $n$-C$_4$H$_{10}$ \\
+15) 3-MP $\rightarrow$ 2,3-DMB  & 39) $n$-C$_5$H$_{12}$ + H$_2$ $\rightarrow$ C$_3$H$_8$ + C$_2$H$_6$ \\
+16) 2,3-DMB $\rightarrow$ 3-MP  & 40) $n$-C$_5$H$_{12}$ + H$_2$ $\rightarrow$ $n$-C$_4$H$_{10}$ + CH$_4$ \\
+17) 2,2-DMB $\rightarrow$ 2,3-DMB  & 41) $iso$-C$_5$H$_{12}$ + H$_2$ $\rightarrow$ $iso$-C$_4$H$_{10}$ + CH$_4$ \\
+18) 2,3-DMB $\rightarrow$ 2,2-DMB & 42) $n$-C$_6$H$_{14}$ + H$_2$ $\rightarrow$ 2C$_3$H$_8$ \\
+19) B + 3H$_2$ $\rightarrow$ CH*  & 43) $n$-C$_6$H$_{14}$ + H$_2$ $\rightarrow$ $n$-C$_5$H$_{12}$ + CH$_4$ \\
+20) CH* $\rightarrow$ B + 3H$_2$ & 44) $n$-C$_6$H$_{14}$ + H$_2$ $\rightarrow$ $n$-C$_4$H$_{10}$ + C$_2$H$_6$ \\
+21) B + 3H$_2$ $\rightarrow$ MCP  & 45) 2-MP + H$_2$ $\rightarrow$ $iso$-C$4$H$_{10}$ + C$_2$H$_6$ \\
+22) MCP $\rightarrow$ B + 3H$_2$  & 46) 3-MP + H$_2$ $\rightarrow$ $iso$-C$_5$H$_{12}$ + CH$_4$ \\
+23) CH* $\rightarrow$ MCP & 47) 2,2-DMB + H$_2$ $\rightarrow$ $iso$-C$_5$H$_{12}$ + CH$_4$ \\
+24) MCP $\rightarrow$ CH*  & 48) 2,3-DMB + H$_2$ $\rightarrow$ $iso$-C$_5$H$_{12}$ + CH$_4$ \\
 \bottomrule
@@
 
-The introduction should briefly place the study in a broad context and highlight why it is important.
-It should define the purpose of the work and its significance.
-The current state of the research field should be reviewed carefully and key publications cited.
-Please highlight controversial and diverging hypotheses when necessary.
-Finally, briefly mention the main aim of the work and highlight the principal conclusions.
-As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
-Citing a journal paper \cite{ref-journal}.
-Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}.
-Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
+Mathematical model of the chemical process is a non-linear system of differential equations \cite{Koledina2016KineticsAM,enik2021,logan}:
+\begin{linenomath}
+\begin{equation}\label{eq_concentrations}
+\dfrac{dx_i}{d\tau} = \sum_{j=1}^J \nu_{ij} w_j, i = 1,...,I,
+\end{equation}
+\begin{equation}\label{eq_rate}
+w_{j} = k_j \prod_{i=1}^I \left( \dfrac{x_i}{F} \right),
+\end{equation}
+\begin{equation}\label{eq_constant}
+k_j = k^0_j \exp \left( -\dfrac{E_j}{RT} \right)
+\end{equation}
+\end{linenomath}
+with initial values: at $\tau = 0$, $x_i(0) = x_i^0$, where $\nu_{ij}$ are stoichiometric coefficients of stages of chemical reaction; $J$ is number of stages ($J = 48$), $x_i$ is molar flow rate of the $i$-th component of reaction, kmol/h; $I$ is the number of components ($I = 17$); $w_j$ is the rate of the $j$-th stage of reaction, kmol/(h$\cdot$kg cat.), $E_j$ is activation energy of reaction stage, J/mol; $R$ is the universal gas constant ($R$=8,31 J/(mol$\cdot$K), $T$ is temperature, K; $k_j$ are pre-exponentional factors, kmol/(h$\cdot$kg cat.); $\tau$ is conditioned contact time, (kg cat.); $F$ is the total molar flow rate, kmol/h.
 
-%Часть ННГУ - начало
-\section{Parallel Algorithm for Solving Global Optimization Problems }\label{Sec_GSA}
+The mathematical model involves such parameters as the raw material flow rate $F$ in the reactor and temperature $T$, but these two parameters also depend on $\tau$, therefore it is necessary to write equations on them, because technological process is non-isothermal:
+\begin{linenomath}
+\begin{equation}\label{eq_temper}
+\dfrac{dT}{d\tau} = \dfrac{\sum_{i=1}^I\left( \dfrac{dx_i}{d\tau} \cdot \Delta H^i_{f,T} \right) }{\sum_i^J \left(x_i C^T_{p_i} \right) },
+\end{equation}\label{eq_flow}
+\begin{equation}
+\dfrac{dF}{d\tau} = \sum_{i=1}^I  \dfrac{x_i}{d\tau}
+\end{equation}
+\end{linenomath}
+with initial condition: at $\tau = 0$, $T(0) = T^0$, where $T$ is temperature, K, $\Delta H^i_{f,T}$ is enthalpy of the $i$-th component formation at temperature $T$, J/mol, $C^T_{p_i}$ is specific heat capacity of the $i$-th component at temperature $T$, (J/mol$\cdot$K), $F$ is mole discharge of the flow, kmol/h.
+
+Thus, when solving the inverse kinetic problem, the direct problem is repeatedly solved, which includes the numerical integration of equations \eqref{eq_concentrations}, \eqref{eq_temper} and \eqref{eq_flow}.
+After finding the numerical values of the molar flow rate $x_i$, temperature $T$ and total molar flow rate $F$, these values are compared with the corresponding industrial data obtained during the three days of operation of the catalytic isomerization unit \cite{enik2021}, which is a minimization task:
+
+\begin{linenomath}
+\begin{equation}
+FF = \dfrac{\sum_{i=1}^I \left| \dfrac{y_i^{exp} - y_i^{calc}}{y_i^{exp}} \right|}{I} \rightarrow \min,
+\end{equation}
+\end{linenomath}
+
+where I is the number of observed parameters (I=19, of which 17 parameters are $x_i$, one parameter is $T$ and one parameter is $F$),  $y_{i}^{exp}$ are industrual data, $y_{i}^{calc}$ are numerical values of concentrations of reaction substances obtained by solving a direct kinetic problem.
+As a result of solving the inverse problem, the kinetic parameters $k^0_j$ and $E_j$ included in the Arrhenius equation \eqref{eq_rate} will be obtained.
+Since the number of reaction stages is 48, the kinetic model will contain 96 kinetic parameters (activation energy $E_j$ and pre-exponential multiplier $k^0_j$ for each stage), however, the authors decided to fix some of the parameters as already known for chemical reasons, so the number of optimized parameters was reduced to 24.
+
+\section{Parallel Algorithm for Solving Global Optimization Problems }\label{sec_GSA}
 
@@
 
-As mentioned above, from a mathematical point of view, the inverse problem of chemical kinetics can be considered a Lipschitz global optimization problem. 
+As stated above, from a mathematical point of view, the inverse problem of chemical kinetics can be considered a Lipschitz global optimization problem. 
 In general, the problem of the specified class can be formulated as follows:
@@
 
-The function $\phi(y)$ is assumed to be multiextremal and is given in the form of some computational algorithm.
-The input of this algorithm is supplied with a parameter vector, and the output is the corresponding value of the function.
+The criterion $\phi(y)$ is assumed to be multiextremal and is given in the form of some computational algorithm.
+The input of this algorithm is supplied with a parameter vector, and the output is the corresponding function value.
 Such functions are often referred to as ``black-box'' functions.
@@
 
-Known examples of such methods are the non-uniform space covering method \cite{Evtushenko2013}, the simplicial partitions method \cite{PaulaviciusZilinskas2014}, the diagonal partitions method \cite{Sergeyev2017}.
+Known examples of such methods are DIvide RECTangles (DIRECT) algorithm \cite{Jones1993}, the simplicial partitions algorithm \cite{PaulaviciusZilinskas2014}, the diagonal partitions algorithm \cite{Sergeyev2017}.
 These techniques are based on adaptive partitioning of the search domain into a finite number of subdomains at each iteration.
@@
 A fundamentally different approach to solving a multidimensional problem (\ref{problemN}) is its reduction to one or more one-dimensional problems followed by the use of efficient methods of one-dimensional optimization.
-Such reduction may be accomplished, for example, by a nested optimization scheme \cite{Grishagin2018} or by Peano-Hilbert curves \cite{Barkalov2018}.
+Such reduction may be accomplished, for example, by a recursive optimization scheme \cite{Grishagin2018} or by Peano curves \cite{Barkalov2018}.
 The latter approach is used in this work.
 
-Using a continuous single-valued mapping (Peano-Hilbert curve) $y(x)$ of the real axis segment $[0,1]$ onto hyperinterval $D$ from (\ref{D}), a multidimensional problem (\ref{problemN}) can be reduced to a one-dimensional one
+Using a continuous single-valued correspondence (Peano curve) $y(x)$ of the real axis segment $[0,1]$ onto hyperinterval $D$ from (\ref{D}), a multidimensional problem (\ref{problemN}) can be reduced to a one-dimensional one
 \[
@@
 \]
-where the H{\"o}lder constant $H$ is associated with the Lipschitz constant $L$ by the relationship $ H=2 L \sqrt{N+3}$.
+where the H{\"o}lder constant $H$ is associated with the Lipschitz constant $L$ by the formula $ H=2 L \sqrt{N+3}$.
 
@@
 
-Thus, under this approach, the search trial at some point $x^k\in[0,1]$ will include the construction of the image $y^k=y(x^k)$ first, and only then the calculation of the value of the function $ z^k = \phi(y^k)$.
+Thus, under this approach, the search trial at some point $x^k\in[0,1]$ will consist of constructing the image $y^k=y(x^k)$ first, and only then of calculating the function value $ z^k = \phi(y^k)$.
 We will call the three values $\{x^k,y ^k,z^k\}$ the trial result.
 
-%поправить название
 \subsection{Asynchronous parallel algorithm used to determine the global minimum of a function}
 
-The parallel algorithm used in this study is based on the information and statistical approach to the development of global optimization methods, the theoretical foundations of which are set out in \cite{Strongin2000}.
+The parallel algorithm used in this study is based on the information-statistical approach to the development of global optimization methods, the theoretical foundations of which are set out in \cite{Strongin2000}.
 
@@
 
-Let us assume that at each iteration, the master-process calculates one point of the new trial and transfers it for testing to the worker-process.
-At the same time, a trial performed by a worker is a much more labor-intensive operation than selection of a new point by a master, which eliminates the downtime of worker-processes.
-In this case (unlike synchronous parallel algorithms), the total number of trials performed by each worker-process  will depend on the labor intensity of the specific trial and cannot be estimated in advance.
+Let us assume that at each iteration, the master calculates one new trial  point and transfers it for processing to the worker.
+At the same time, a trial performed by a worker is a much more labor-intensive operation than selection of a new point by a master, which eliminates the worker downtime.
+In this case, the total number of trials performed by each worker  will depend on the labor intensity of the specific trial and cannot be estimated in advance.
 
@@
 The initial iteration is carried out according to a special rule.
-The master process (let us assume that this is the process No 0) initiates  $p$ parallel trials at $p$ different points of the search domain, two of which are boundary points, and the rest are arbitrary internal points (for example, uniformly distributed).
+The master process (let us assume that this is the process No 0) initiates  $p$ parallel trials at $p$ different points of the search region, two of which are boundary points, and the rest are arbitrary internal points (for example, uniformly distributed).
 Thus, the initial trials are carried out at the points $\{y(x^1), y(x^2), ...,y(x^p)\}$, where $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.
@@
 Let one of the processes (without limiting the commonality, we can assume that this is the process No 1), at some point in time complete the trial at the point $y(x^{k+1})$.
-The remaining processes are at the stage of conducting their trials, i.e. at these points multiple trials have already started, but have not yet been completed.
+The remaining processes are at the stage of conducting their trials, i.e. at these points multiple trials have already initiated, but have not yet been finalized.
 
@@
 where $z_i=\phi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$.
-The $z_i$ values at $x_i \in I_k$ are undefined because the trials at $x_i \in I_k$ are not yet complete.
 If the value $M$ is 0, assign $M=1$.
+The values $z_i=\phi(y(x_i))$ at $x_i \in I_k$ are undefined because the trials at $x_i \in I_k$ are not finished yet.
 
@@
 
-Immediately after calculating the point $x^{k+p+1}$ of the next trial, the master-process adds it to the set $I_k$ and forwards it to the worker-process, which initiates the trial at this point.
+Immediately after calculating the point $x^{k+p+1}$ of the next trial, the master appends it to the set $I_k$ and forwards it to the worker, which starts the trial at this point.
 
-The master process terminates the algorithm if either the condition $\Delta_{t}<\varepsilon$ ($t$ is number of the interval with the maximum characteristic from Step 4) or the condition $k>K_{max}$ ($k$ is the number of trials performed) is met.
-The real number $\varepsilon>0$ and the integer $K_{max}>0$ are the algorithm parameters which correspond to the accuracy of the solution search and the maximum number of trials.
+The master stops the algorithm if either the condition $\Delta_{t}<\varepsilon$ (where $t$ is number of the interval with the maximum characteristic from Step 4) or the condition $k>K_{max}$ (where $k$ is the number of trials performed) is met.
+The real number $\varepsilon>0$ and the integer $K_{max} \geq 1$ are the algorithm parameters which correspond to the accuracy of the solution search and the maximum number of trials.
 
@@
 
-\section{Numerical Experiments}
+\section{Numerical Experiments}\label{sec_numexper}
 
@@
 To solve the optimization problem, a combination of the parallel global search algorithm and a local optimization method was used.
-In the first stage of the search, the parallel asynchronous algorithm described in Section \ref{Sec_GSA} was used.
+In the first stage of the search, the parallel asynchronous algorithm described in Section \ref{sec_GSA} was used.
 101 processes were involved in the launch, of which one process was a master and the remaining 100 were workers (calculators).
@@
 The total time for solving the problem was 8109 seconds, an optimum point was found with the objective function value $\phi^* = 0.002115$.
-
-%Часть ННГУ - конец
+Thus, using the developed algorithm, the inverse problem of chemical kinetics was solved and the kinetic parameters of the reaction were found, which are given in the table \ref{tab_kinetics}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
+\section{Discussion}\label{sec_discussion}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses.
+%Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses.
 The findings and their implications should be discussed in the broadest context possible.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
+After the development of the kinetic reaction model, it is necessary to check the adequacy of the model, this procedure can be carried out using a direct kinetic problem.
+Substituting the found kinetic values $k_j^0$ and $E_j$ into the system \eqref{eq_concentrations} -- \eqref{eq_flow}, the direct problem of chemical kinetics was solved and the concentrations of reaction substances $x_i$ and temperature $T$ were found (Fig.
+\ref{fig:res1} -- \ref{fig:res2}).
 
-This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
+Figures \ref{fig:res1} -- \ref{fig:res2} show graphs of changes in the concentrations of individual components of the reaction and temperature change in the reactor unit.
+It should be noted that in the figures the vertical lines separating the graphs characterize three consecutive reactors (I, II, III).
+The initial data on the concentrations of the components were known only at the entrance to reactor I and at the exit from reactor III.
+The concentrations of components at the outlet of reactor III are marked with a marker in the figures.
+
+\begin{figure}
+\center
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=0.9\linewidth]{1.eps} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=0.9\linewidth]{2.eps} \\ (b)}
+\end{minipage}
+\caption{Changes in concentrations of (a) $n$-pentane ($n$-C$_5$) and $n$-hexane ($n$-C$_6$), (b) isopentane (i-C5), 2-methylpentane (2-MP), 3-methylpentane (3-MP), 2,2-dimethylbutane(2,2-DMB), 2,3-dimethylbutane (2,3-DMB).}\label{fig:res1}
+\end{figure}
+
+\begin{figure}
+\center
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=0.9\linewidth]{3.eps} \\ (a)}
+\end{minipage}
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=0.9\linewidth]{4.eps} \\ (b)}
+\end{minipage}
+\caption{Changes in (a) concentrations of benzene, cyclohexane (CH*), methyl cyclopentane (MCP), cyclopentane (CP), (b) temperature.}\label{fig:res2}
+\end{figure}
+
+Figure \ref{fig:res1} (a) shows that the concentrations of n-pentane and n-hexane are continuously decreasing.
+This is mainly due to the transformation of alkanes of normal (linear) structure into the corresponding isomers, which is reflected in Figure \ref{fig:res1} (b).
+With the transformation of n-pentane, everything is natural, since only one isomer is known.
+And in the case of n-hexane, everything is somewhat different: first, the conversion of n-hexane is carried out into isomers 2-methylpentane and 3-methylpentane, and the latter are further converted into 2,2-dimethylbutane and 2,3-dimethylbutane.
+
+According to Figure \ref{fig:res2} (a), it can be seen that the concentration of cyclopentane in reactor I gradually decreases to zero.
+The decrease in cyclopentane concentration can be explained by the hydrogenation reaction to $n$-pentane.
+According to Figure \ref{fig:res1} (a), this is probably the reason for the increase in the concentration of $n$-pentane at the beginning of reactor I.
+Also, this figure shows a sharp decrease in the concentration of benzene, presumably due to hydrogenation reactions.
+
+Figure \ref{fig:res2} (b) shows a graph of temperature changes in reactors I-III.
+According to the technological scheme, refrigeration equipment for cooling the reaction mixture is installed after reactor II.
+A gradual increase in temperature in reactors III here indicates the presence of isomerization reactions that occur with a slight exothermic effect.
+
+It is worth noting that the composition of the products at the outlet of the last reactor and the temperature differences in all reactors are close to industrial data.
+Thus, these facts may indicate the adequacy of the model we have developed.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Patents}
-
-%This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}
-
-% Only for the journal Methods and Protocols:
-% If you wish to submit a video article, please do so with any other supplementary material.
-% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
-A supporting video article is available at doi: link.}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``Conceptualization, X.X.
-and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
-All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
-Authorship must be limited to those who have contributed substantially to the work~reported.}
+\authorcontributions{Conceptualization, K.B.
+and I.G.; methodology, K.B.
+and I.G.; software, E.K., I.L.
+and L.E.; validation, A.F.;  resources, A.F.;  writing---original draft preparation, K.B.
+and A.F.; writing---review and editing, R.F.
+and L.E.; visualization, R.F.; supervision, I.G..
+All authors have read and agreed to the published version of the manuscript.}
 
-\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
-Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
+\funding{This research was funded by RFBR, project No.\,19-37-60014 (L.E.; mathematical model), and by the Russian Science Foundation, project No.\,21-11-00204 (K.B, I.L., E.K.; parallel algorithm)}
 
@@
 
-\dataavailability{In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study.
-Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.
-If the study did not report any data, you might add ``Not applicable'' here.}
+\dataavailability{Not applicable.} 
+\conflictsofinterest{The authors declare no conflict of interest.}
 
-%\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
-This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+\abbreviations{Abbreviations}{
+The following abbreviations are used in this manuscript:\\
 
-\conflictsofinterest{The authors declare no conflict of interest.}
+\noindent 
+\begin{tabular}{@{}ll}
+2-MP & 2-methylpentane\\
+3-MP & 3-methylpentane\\
+2,2-DMB  & 2,2-dimethylbutane\\
+2,3-DMB & 2,3-dimethylbutane\\
+CH* & cyclohexane\\
+MCP & methylcyclopentane\\
+B & benzene\\
+CP & cyclopentane
+\end{tabular}
+}
 
@@
 %% Optional
-\sampleavailability{Samples of the compounds ... are available from the authors.}
+\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixstart
+\appendix
+\section[\appendixname~\thesection]{}
+%\subsection[\appendixname~\thesubsection]{}
+The table \ref{tab_kinetics} is a numerically calculated kinetic parameters of a chemical reaction.
 
-%% Only for journal Encyclopedia
-%\entrylink{The Link to this entry published on the encyclopedia platform.}
-
-%\abbreviations{Abbreviations}{
-%The following abbreviations are used in this manuscript:\\
-%
-%\noindent 
-%\begin{tabular}{@{}ll}
-%MDPI & Multidisciplinary Digital Publishing Institute\\
-%DOAJ & Directory of open access journals\\
-%TLA & Three letter acronym\\
-%LD & Linear dichroism
-%\end{tabular}
-%}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional
+\begin{table}[H] 
+\caption{Kinetic parameters of the catalytic isomerisation reaction of the pentane-hexane fraction, according to the scheme of chemical transformations (Table \ref{tab_chemistry}).\label{tab_kinetics}}
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+\begin{tabularx}{\textwidth}{CCCCCC}
+\toprule
+\textbf{No}	& \textbf{$E$, kcal/mol} & \textbf{$\log k_0$} & \textbf{No}	& \textbf{$E$, kcal/mol} & \textbf{$\log k_0$} \\
+\midrule
+1  & 149.320 & 18.492 & 25 & 141.046 & 5.799 \\
+2  & 129.372 & 15.604 & 26 & 80.093 & 4.401 \\
+3  & 108.439 & 9.286 & 27 & 177.947 & 3.694 \\
+4  & 87.168 & 9.002 & 28 & 92.282 & 10.670  \\
+5  & 123.830 & 14.203 & 29 & 159.368 & 16.169 \\
+6  & 162.603 & 17.997 & 30 & 97.215 & 9.612 \\
+7 & 161.429 & 17.114 & 31 & 188.306 & 7.194  \\
+8 & 142.261 & 12.667  & 32 & 111.789 & 11.467 \\
+9 & 127.715 & 12.130 & 33 & 124.370 & 12.764 \\
+10 & 96.409 & 11.318 & 34 & 101.952 & 6.352 \\
+11 & 64.223 & 7.214 & 35 & 131.079 & 16.427  \\
+12 & 94.973	& 8.388 & 36 & 177.771 & 11.250  \\
+13 & 174.387 & 7.600 & 37 & 72.632 & 15.123 \\
+14 & 161.502 & 18.146 & 38 & 67.167 & 14.013 \\
+15 & 98.296 & 11.305 & 39 & 230.976 & 12.798 \\
+16 & 104.987 & 11.499 & 40 & 214.539 & 13.532  \\
+17 & 42.264 & 3.622 & 41 & 291.967 & 26.096  \\
+18 & 45.070 & 5.530 & 42 & 185.864 & 20.281 \\
+19 & 162.126 & 21.057 & 43 & 134.411 & 14.229 \\
+20 & 231.848 & 28.391 & 44 & 82.0120 & 8.262 \\
+21 & 245.107 & 28.488 & 45 & 388.331 & 17.056 \\
+22 & 204.298 & 23.176 & 46 & 286.722 & 18.898 \\
+23 & 53.493 & 5.911 & 47 & 356.242 & 5.148  \\
+24 & 252.728 & 1.600 & 48 & 355.642 & 9.141  \\
+\bottomrule
+\end{tabularx}
+\end{table}
 
@@
 %=====================================
-ё\bibliography{bibliography}
+\bibliography{bibliography.bib}
 
@@
 %=====================================
+%\begin{thebibliography}{999}
+
+% Reference 1
+%\bibitem[Author1(year)]{ref-journal}
+%Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
+% Reference 2
+%\bibitem[Author2(year)]{ref-book1}
+%Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor 1, F., Editor 2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
+% Reference 3
+%\bibitem[Author3(year)]{ref-book2}
+%Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
+% Reference 4
+%\bibitem[Author4(year)]{ref-unpublish}
+%Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} year, \textit{phrase indicating stage of publication (submitted; accepted; in press)}.
+% Reference 5
+%\bibitem[Author5(year)]{ref-communication}
+%Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
+% Reference 6
+%\bibitem[Author6(year)]{ref-proceeding}
+%Author 1, A.B.; Author 2, C.D.; Author 3, E.F.
+Title of presentation.
+In Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference (Day Month Year); Abstract Number (optional), Pagination (optional).
+% Reference 7
+%\bibitem[Author7(year)]{ref-thesis}
+%Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
+% Reference 8
+%\bibitem[Author8(year)]{ref-url}
+%Title of Site. Available online: URL (accessed on Day Month Year).
+%\end{thebibliography}
 

### 2022-10-04 2022/RSD/paper.tex
submitted to Springer

@@
 \usepackage{multirow} 
+\usepackage{marvosym}
+
 %
@@
 \author{
-Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
+Konstantin Barkalov$^\textrm{\Letter}$ \orcidID{0000-0001-5273-2471} \and
 Ilya Lebedev \orcidID{0000-0002-8736-0652} \and

### 2022-10-11 2022/NumMethOneAPI/NMP_authorGuidelinesRu_2022.tex
Обновленный текст статьи

@@
 
-\Citation{К.А. Баркалов, И.Г. Лебедев, Я.В. Кольтюшкина
+\Citation{Баркалов К.А., Лебедев И.Г., Силенко Я.В.
     Решение трудоемких задач многомерной глобальной оптимизации с использованием набора инструментов Intel oneAPI~//
@@
     doi 10.26089/NumMet.v??r???.}
-    {K.A. Barkalov, I.G. Lebedev, Ya.V. Koltyushkina,
+    {K.~A.~Barkalov, I.~G.~Lebedev, Ya.~V.~Silenko,
     Solving time-consuming multidimensional global optimization problems with Intel OneApi tools 
@@
 
-\Author{	К.А. Баркалов}{K.A. Barkalov}
-\FullName{Баркалов Константин Александрович}{Barkalov Konstantin Aleksandrovich}
+\Author{К.~А.~Баркалов}{Konstantin~A.~Barkalov}
+\FullName{Баркалов Константин Александрович}{Konstantin~A.~Barkalov}
 % Место работы 1
@@
 
-\Author{ И.Г. Лебедев}{ I.G. Lebedev}
-\FullName{Лебедев Илья Генадьевич}{ Lebedev Ilya Genadievich}
+\Author{И.~Г.~Лебедев}{Ilya~G.~Lebedev}
+\FullName{Лебедев Илья Генадьевич}{Ilya~G.~Lebedev}
 % Место работы 1
@@
 
-\Author{ Я.В. Кольтюшкина}{ Ya.V. Koltyushkina}
-\FullName{ Кольтюшкина Янина Вадимовна}{ Koltyushkina Yanina Vadimovna}
+\Author{Я.~В.~Силенко}{Yanina~V.~Silenko}
+\FullName{Силенко Янина Вадимовна}{Yanina~V.~Silenko}
 % Место работы 1
@@
 \label{task}
-\varphi(y^*)=\min\{\varphi(y):y\in D\}
+\varphi(y^*)=\min\{\varphi(y):y\in D\},
 \end{equation}
@@
 \label{lip}
-|\varphi(y_1)-\phi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
+|\varphi(y_1)-\phi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty.
 \end{equation}
@@
 \label{cube}
-\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
+\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1}, 1\leqslant i\leqslant N\rbrace=\{y(x): 0\leqslant x\leqslant 1\},
 \end{equation}
@@
 \label{oneDimTask}
-\varphi(y^*)=\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}
+\varphi(y^*)=\varphi(y(x^*))=\min\{\varphi(y(x)): x\in [0,1]\}.
 \end{equation}
@@
 \label{holder}
-|\phi(y(x_1))-\phi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
+|\phi(y(x_1))-\phi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}}, 
+ x_1,x_2\in[0,1],
 \end{equation}
@@
 \begin{equation}
-H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}
+H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}.
 \end{equation}
@@
 Алгоритм глобального поиска относится к классу характеристических алгоритмов оптимизации, хорошо масштабируется и отлично подходит для работы на параллельных вычислительных кластерах.
+
+\section{Параллельный алгоритм глобального поиска}
+
 В процессе своей работы алгоритм порождает последовательность точек $x^k$, в каждой из которых проходит вычисление значения минимизируемой функции $z^k=f(x^k)$.
@@
 \label{agp1_sort}
-	0=x_0<\ x_1<\ ...\ <x_{k+1}=1 
+	0=x_0<\ x_1<\ ...\ <x_{k+1}=1,
 	\end{equation}
@@
 \label{agp3_R1}
-R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M}
+R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
 \end{equation}
@@
 \label{agp3_Ri}
-R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1
+R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1,
 \end{equation}
@@
 \label{agp4_R_sort}
-	R\left(t_1\right)\geq\ R\left(t_2\right)\geq...\geq\ R\left(t_k\right)\geq\ R(t_{k+1})\
+	R\left(t_1\right)\geq\ R\left(t_2\right)\geq...\geq\ R\left(t_k\right)\geq\ R(t_{k+1}),\ 
 \end{equation}
@@
 \label{agp5_x1}
-	x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\ t_j=1,\ t_j=k+1
+	x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\ t_j=1,\ t_j=k+1,
 \end{equation}
@@
 \label{agp4_xi}	
-	x^{k+1}=\frac{x_{t_j}+x_{t_j-1}}{2}-sign\left(z_{t_j}-z_{t_j-1}\right)\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N,\ 1<t_j<k+1
+	x^{k+1}=\frac{x_{t_j}+x_{t_j-1}}{2}-sign\left(z_{t_j}-z_{t_j-1}\right)\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N,\ 1<t_j<k+1.
 \end{equation}
@@
 \begin{equation}
-f_k^*=\min_{1\leqslant i \leqslant k}f(x_i), x_k^*=arg \min_{1\leqslant i \leqslant k}f(x_i)
+f_k^*=\min_{1\leqslant i \leqslant k}f(x_i), x_k^*=arg \min_{1\leqslant i \leqslant k}f(x_i).
 \end{equation}
+
+\section{Теоретические оценки ускорения параллельного алгоритма}
+Опишем (по книге [\ref{rfa:rulit:lit2}]) теоретические свойства параллельного алгоритма.
+В решаемых задачах время выполнения одного испытания существенно превышает время обработки ее результата.
+Таким образом, ускорение по испытаниям будет:
+
+\begin{equation} \label{par_trl_ref}
+s(p) = \frac{n(1) \cdot p}{n(p)}
+\end{equation}
+
+$s(p)$ является ключевой характеристикой эффективности параллельного алгоритма глобального поиска.
+Где $n(1)$ — количество испытаний, выполненных последовательным алгоритмом, а $n(p)$ — количество испытаний, выполненных параллельным алгоритмом с использованием $p$ параллельных устройств.
+
+Очевидно, что количество испытаний $n(p)$ для алгоритмов с разной степенью распараллеливания $p$ будут отличаться друг от друга.
+Действительно, последовательный алгоритм имеет полную информацию, полученную на предыдущих $k$ итерациях, при выборе точки $x^{k+1}$ следующего $(k+1)-ого$ испытания.
+Параллельный алгоритм выбирает не одну, а $p$ точек $x^{k+j}$, $1 \leq j \leq p,$ на основе той же информации.
+Это означает, что выбор точек $x^{k+j}$ производится без информации о результатах испытаний в точках $x^{k+i}$, $1 \leq i < j$.
+Только первая точка $x^{k+1}$ будет соответствовать точке, выбранной последовательным алгоритмом.
+Точки других испытаний, как правило, могут не совпадать с генерируемыми последовательным алгоритмом.
+Поэтому будем называть такие испытания излишними, а количество
+\begin{displaymath}
+\lambda(p) = \left\{ \begin{array}{ll}
+                (n(p) - n(1)) / n(p), & \textrm{$n(p) > n(1)$}\\
+                0, & \textrm{$n(p) \leq n(1)$}
+  \end{array} \right.
+\end{displaymath}
+будем называть избыточностью метода.
+
+Следующие утверждения из [\ref{rfa:rulit:lit2}] определяют степень распараллеливания $p$, которая соответствует неизбыточному (т.е. с нулевой избыточностью) распараллеливанию.
+
+Обозначим серии испытаний, генерируемые последовательным алгоритмом и параллельным при решении одной и той же задачи с $\varepsilon=0$ в условиях остановки, как $\{x^k\}$ и $\{y^m \}$ соответственно.
+
+\textbf{Теорема}. Пусть $x^*$ — точка глобального минимума, а $x^{\prime}$ — точка локального минимума функции $f(x)$, и пусть выполнены следующие условия:
+    \begin{enumerate}
+        \item Выполняется неравенство
+            \begin{equation} \label{first_s_ref}
+                f(x^{\prime}) - f(x^*) \leq \delta, \delta > 0.
+            \end{equation}
+        \item Первые $q(l)$ испытаний последовательного и параллельного алгоритмов совпадают, т.е.
+            \begin{equation} \label{second_s_ref}
+                \{x^1,...,x^{q(l)}\} = \{y^1,...,y^{q(l)}\}
+            \end{equation}
+        где
+            \begin{equation} \label{third_s_ref}
+                \{x^1,...,x^{q(l)}\} \subset \{x^k\}, \{y^1,...,y^{q(l)}\}\subset \{y^m\}.
+            \end{equation}
+        \item Существует точка $t^n \in \{y^m\}$, $n < q(l)$, такая что $x^{\prime} \leq y^n \leq x^*$ или $x^* \leq y^n \leq x^{\prime}$.
+        \item Для величины $M$ из (\ref{agp2_mu}) выполняется неравенство 
+            \begin{equation} \label{fourth_s_ref}
+                M > 2^{2 - 1/N} H,
+            \end{equation}
+        где $H$ — постоянная Гёльдера целевой одномерной функции.
+    \end{enumerate}
+
+Тогда параллельный алгоритм при $p=2$ будет неизбыточным (т.е. $s(2)=2$, $\lambda(2)=0$) до тех пор, пока выполняется условие
+
+\begin{equation} \label{p_two_ref}
+    (x_{t_j} - x_{t_j - 1})^{1/N} > \frac{4\delta}{M - 2^{2 - 1/N} H},
+\end{equation}
+
+где $t_j$ определяется в соответствии с шагом 4 алгоритма.
+
+\textbf{Следствие}.
+
+Пусть целевая функция $f(x)$ имеет $Q$  локальных минимумов $\{x_1^{\prime},...,x_Q^{\prime}\}$, для которых выполняется условие \eqref{first_s_ref} и пусть существуют точки испытания $y^{n_i}$, $1 \leq i \leq Q,$ такие, что
+\begin{gather} 
+    y^{n_i} \in \{y^1,...y^{q(l)}\}, \nonumber \\ 
+    \alpha_i \leq y^{n_i} \leq \alpha_{i+1}, \; \alpha_i, \alpha_{i+1} \in \{x^*, x_1^{\prime},...,x_Q^{\prime}\}, 1 \leq i \leq Q. \nonumber
+\end{gather}
+Тогда при выполнении условий теоремы, параллельный алгоритм со степенью распараллеливания $Q+1$ будет неизбыточным (т.е.
+$s(Q+1)=Q+1, \; \lambda(Q+1) =0$) до тех пор, пока не будет выполнено условие \eqref{p_two_ref}.
+
+Следствие из теоремы играет особую роль при решении многомерных задач, сводящихся к одномерным с помощью отображения $y(x)$.
+Отображение $y(x)$, являясь аппроксимацией кривой Пеано, имеет эффект "расщепления" точки глобального минимума $y^* \in D$ на несколько (до $2^N$) прообразов в интервале [0,1].
+Применяя параллельный алгоритм для минимизации редуцированной одномерной задачи, можно ожидать нулевую избыточность при степени распараллеливания до $2^N+1$.
 
@@
 \begin{displaymath}
-\underset{{{i}_{1}}=0}{\overset{M-1}{\mathop \sum }}\,\underset{{{i}_{2=0}}}{\overset{M-1}{\mathop \sum }}\,..\underset{{{i}_{N}}=0}{\overset{M-1}{\mathop \sum }}\,\left( \left( \underset{j=1}{\overset{2N}{\mathop \prod }}\,{{h}_{j}} \right)\psi \left( {{y}_{1}},{{y}_{2}},~..,{{y}_{N}},~{{y}_{{{\left( N+1 \right)}_{{{i}_{1}}}}}},{{y}_{{{\left( N+2 \right)}_{{{i}_{2}}}}}}~,~..,{{y}_{{{\left( 2N \right)}_{{{i}_{N}}}}}} \right) \right)
+\underset{{{i}_{1}}=0}{\overset{M-1}{\mathop \sum }}\,\underset{{{i}_{2=0}}}{\overset{M-1}{\mathop \sum }}\,..\underset{{{i}_{N}}=0}{\overset{M-1}{\mathop \sum }}\,\left( \left( \underset{j=1}{\overset{2N}{\mathop \prod }}\,{{h}_{j}} \right)\psi \left( {{y}_{1}},{{y}_{2}},~..,{{y}_{N}},~{{y}_{{{\left( N+1 \right)}_{{{i}_{1}}}}}},{{y}_{{{\left( N+2 \right)}_{{{i}_{2}}}}}}~,~..,{{y}_{{{\left( 2N \right)}_{{{i}_{N}}}}}} \right) \right),
 \end{displaymath}
 \begin{displaymath}
-D=\left\{ y~\in ~{{R}^{2N}}:~{{a}_{i}}\le ~~{{y}_{i}}\le ~~{{b}_{i}},~1\le i\le 2N \right\}
+D=\left\{ y~\in ~{{R}^{2N}}:~{{a}_{i}}\le ~~{{y}_{i}}\le ~~{{b}_{i}},~1\le i\le 2N \right\},
 \end{displaymath}
 \begin{equation}
-{{y}_{{{\left( N+k \right)}_{{{i}_{k}}}}}}={{a}_{k}}+{{i}_{k}}*{{h}_{k}}+\frac{{{h}_{k}}}{2}
+{{y}_{{{\left( N+k \right)}_{{{i}_{k}}}}}}={{a}_{k}}+{{i}_{k}}*{{h}_{k}}+\frac{{{h}_{k}}}{2},
 \end{equation}
@@
 \label{Rastrigin}
-\psi \left( {{y}_{1}},\ldots ,{{y}_{2N}}~ \right)=~\underset{i=1}{\overset{2N}{\mathop \sum }}\,(y_{i}^{2}-10\cos \left( 2\pi {{y}_{i}} \right)+10)
+\psi \left( {{y}_{1}},\ldots ,{{y}_{2N}}~ \right)=~\underset{i=1}{\overset{2N}{\mathop \sum }}\,(y_{i}^{2}-10\cos \left( 2\pi {{y}_{i}} \right)+10),
 \end{equation}
@@
 \begin{displaymath}
--2.2<y\_i<1.8~,~1\le i\le 2N
+-2.2<y\_i<1.8~,~1\le i\le 2N.
 \end{displaymath}
@@
 \rlitem{rfa:rulit:lit1}{
-\textit{Sergeyev, Ya.D.} Introduction to global optimization exploiting space-filling curves / Ya.D. Sergeyev, R.G. Strongin, D. Lera – Springer, 2013. 125 p.}
+\textit{Sergeyev Ya.D., Strongin R.G, Lera D.} Introduction to global optimization exploiting space-filling curves. Springer, 2013. 125 p.}
 
 \rlitem{rfa:rulit:lit2}{
-\textit{Strongin, R.G.} Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms / R.G. Strongin, Ya.D. Sergeyev – Kluwer Academic Publishers, 2000. 704 p.}
+\textit{Strongin R.G., Sergeyev Ya.D.} Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, 2000. 704 p.}
 
 \rlitem{rfa:rulit:lit3}{
-\textit{Стронгин, Р.Г.} Параллельные вычисления в задачах глобальной оптимизации / Р.Г. Стронгин, В.П. Гергель, В.А. Гришагин, К.А. Баркалов – М.: Издательство Московского университета, 2013. 280 с.}
+\textit{Стронгин Р.Г., Гергель В.А., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013. 280 с.}
 
 \rlitem{rfa:rulit:lit4}{
-\textit{Pinter (Ed.), J.D.} Global Optimization: Scientific and Engineering Case Studies / J.D. Pinter – Springer, 2006. 546 p.}
+\textit{Pinter (Ed.), J.D.} Global Optimization: Scientific and Engineering Case Studies. Springer, 2006. 546 p.}
 
 \rlitem{rfa:rulit:lit5}{
-\textit{Стронгин, Р.Г.} Параллельные методы решения задач глобальной оптимизации / Р.Г.
-Стронгин, В.П.
-Гергель, К.А Баркалов, // Известия высших учебных заведений.
-Приборостроение, 2009.
-Т.
-52.
-№ 10.
-С.
-25 – 33.}
+\textit{Стронгин Р.Г., Гергель В.А., Баркалов К.А.} Параллельные методы решения задач глобальной оптимизации. Известия высших учебных заведений. Приборостроение, 2009. Т. 52. № 10. С. 25 – 33.}
 
 \rlitem{rfa:rulit:lit6}{
-\textit{Захарова, Е.} Обзор методов многомерной оптимизации. / Е.М. Захарова, И.К. Минашина, // Информационные процессы, 2014. Т 14. №3. С. 256 – 274.}
+\textit{Захарова Е.М., Минашина И.К.} Обзор методов многомерной оптимизации. Информационные процессы, 2014. Т 14. №3. С. 256 – 274.}
 
 \rlitem{rfa:rulit:lit7}{
-\textit{Gaviano, M.} Software for generation of classes of test functions with known local and global minima for global optimization/ M.
-Gaviano, D.
-Lera, D.
-E.
-Kvasov, Y.
-D.
-Sergeyev // ACM Transactions on Mathematical Software.
+\textit{Gaviano M., Lera D., Kvasov D.E., Sergeyev Y.D.} Software for generation of classes of test functions with known local and global minima for global optimization.
+ACM Transactions on Mathematical Software.
 – 2003.
@@
 \rlitem{rfa:rulit:lit8}{
-\textit{Сергеев, Я.Д.} Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. – 352 c.}
+\textit{Сергеев Я.Д., Квасов Д.Е.} Диагональные методы глобальной оптимизации.  М.: Физматлит, 2008. – 352 c.}
 
 \rlitem{rfa:rulit:lit9}{
-\textit{Gablonsky, J.M.} A Locally-Biased Form of the DIRECT Algorithm / J.M.  Gablonsky, C.T. Kelley, // Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.}
+\textit{Gablonsky J.M., Kelley C.T.} A Locally-Biased Form of the DIRECT Algorithm. Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.}
 
 \rlitem{rfa:rulit:lit10}{
-\textit{Paulavicius, R.} Parallel branch and bound for global optimization with combination of Lipschitz bounds / R.
-Paulavicius, J.
-Zilinskas and A.
-Grothey // Optimization Methods \& Software,  2011.
+\textit{Paulavicius R., Zilinskas J, Grothey A.} Parallel branch and bound for global optimization with combination of Lipschitz bounds.
+Optimization Methods \& Software,  2011.
 – Vol.
@@
 \rlitem{rfa:rulit:lit11}{
-\textit{Боресков А.А.} Параллельные вычисления на GPU.
-Архитектура и программная модель CUDA / Боресков А.А., Харламов А.А., Марковский Н.Д., Микушин Д.Н., Мортиков Е.В., Мыльцев А.А., Сахарных Н.А., Фролов В.А.
-– М.:  Издательство Московского университета, 2015.
+\textit{Боресков А.А., Харламов А.А., Марковский Н.Д., Микушин Д.Н., Мортиков Е.В., Мыльцев А.А., Сахарных Н.А., Фролов В.А.} Параллельные вычисления на GPU.
+Архитектура и программная модель CUDA.
+М.:  Издательство Московского университета, 2015.
 336 с.}
@@
 \rlitem{rfa:rulit:lit12}{
-\textit{Таненбаум, Э.} Архитектура компьютера / Э. Таненбаум, Т. Остин; перевод с английского Е. Матвеева. — 6-е изд. — Санкт-Петербург [и др.]: Питер, 2014. 816 с.}
+\textit{Таненбаум Э., Остин Т.} Архитектура компьютера / перевод с английского Е. Матвеева. — 6-е изд. — Санкт-Петербург [и др.]: Питер, 2014. 816 с.}
 
 \rlitem{rfa:rulit:lit13}{
-\textit{Комолов Д.А.} Системы автоматизированного проектирования фирмы Altera MAX+Plus II и Quartus II. / Комолов Д.А. Мяльк Р.А. Зобенко А.А. Филиппов А.С., Издательство РадиоСофт, 2002. 355 c.}
+\textit{Комолов Д.А., Мяльк Р.А., Зобенко А.А., Филиппов А.С.} Системы автоматизированного проектирования фирмы Altera MAX+Plus II и Quartus II. Издательство РадиоСофт, 2002. 355 c.}
 
 \rlitem{rfa:rulit:lit14}{
-\textit{Gergel, V.P.} A novel supercomputer software system for solving time-consuming global optimization problems / V.P.
-Gergel, K.A.
-Barkalov A.V.
-Sysoyev // Numerical Algebra, Control and Optimization, 2018.
+\textit{Gergel V.P., Barkalov K.A., Sysoyev A.V.} A novel supercomputer software system for solving time-consuming global optimization problems.
+Numerical Algebra, Control and Optimization, 2018.
 Vol.
@@
 \rlitem{rfa:rulit:lit16}{
-\textit{Сергеев, Я.Д.} Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. 352 c.}
+\textit{Сергеев, Я.Д., Квасов Д.Е.} Диагональные методы глобальной оптимизации. М.: Физматлит, 2008. 352 c.}
 
-\rlitem{rfa:rulit:lit17}{
-\textit{Gaviano, M.} Software for generation of classes of test functions with known local and global minima for global optimization/ M.
+%\rlitem{rfa:rulit:lit17}{
+%\textit{Gaviano, M., Lera D., Kvasov D.E., Sergeyev Y.D.} Software for generation of classes of test functions with known local and global minima for global optimization/ M.
 Gaviano, D.
@@
 
-\elitem{rfa:enlit:COPE1}{
-Guidelines on Good Publication Practice. Committee on Publication Ethics (COPE).
-\url{https://publicationethics.org/files/u7141/1999pdf13.pdf}.
-Cited November 12, 2021.}
+\elitem{rfa:rulit:lit1}{
+\textit{Sergeyev Ya.D., Strongin R.G, Lera D.} Introduction to global optimization exploiting space-filling curves. Springer, 2013. 125 p.}
 
-\elitem{rfa:enlit:Kotelnikov}{
-I.~A.~Kotel'nikov and P.~Z.~Chebotaev, 
-\textit{\LaTeX\ in Russian} 
-(Sibirskiy Khronograph, Novosibirsk, 2004) [in Russian].}
+\elitem{rfa:rulit:lit2}{
+\textit{Strongin R.G., Sergeyev Ya.D.} Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, 2000. 704 p.}
 
-\elitem{rfa:enlit:Samarin}{
-M.~Goossens, F.~Mittelbach, and A.~Samarin, 
-\textit{The \LaTeX~Companion.} 
-(Addision--Wesley Professional, 1994; Mir, Moscow, 1999) 
-[in Russian].}
+\elitem{rfa:rulit:lit3}{
+\textit{Strongin R.G., Gergel V.A., Grishagin V.A., Barkalov K.A.} %Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013. 280 с.
+}
 
-\elitem{rfa:enlit:Lvovski}{
-S.~M.~L'vovsky. \textit{Typesetting and Layout in the \LaTeX~System}
-(MCNMO Publ., Moscow, 2003) 
-[in Russian].}
+\elitem{rfa:rulit:lit4}{
+\textit{Pinter (Ed.), J.D.} Global Optimization: Scientific and Engineering Case Studies. Springer, 2006. 546 p.}
 
-\elitem{rfa:enlit:rot}{
-Ya.~S.~Bugrov and S.~M.~Nikol'sky, 
-\textit{Higher Mathematics: Differential Equations. Multiple Integrals. Ranks} 
-(Phoenix Publ., Rostov-na-Donu, 1998) 
-[in Russian].}
+\elitem{rfa:rulit:lit5}{
+\textit{Strongin R.G., Gergel V.A., Barkalov K.A.} %Параллельные методы решения задач глобальной оптимизации. Известия высших учебных заведений. Приборостроение, 2009. Т. 52. № 10. С. 25 – 33.
+}
 
-\elitem{rfa:enlit:Antonov}{
-A.~S.~Antonov, 
-\textit{Parallel Programming Using OpenMP Technology}
-(Mosk. Gos. Univ., Moscow, 2009) 
-[in Russian].}
+\elitem{rfa:rulit:lit6}{
+\textit{Zakharova E.M., Minashina I.K.} %Обзор методов многомерной оптимизации. Информационные процессы, 2014. Т 14. №3. С. 256 – 274.
+}
 
-\elitem{rfa:enlit:Lapko}{
-O.~Lapko, 
-\textit{The Makecell Package}. 
-\url{https://mirror.truenetwork.ru/CTAN/macros/latex/contrib/makecell/makecell-rus.pdf}.
-Cited November 12, 2021 [in Russian].}
+\elitem{rfa:rulit:lit7}{
+\textit{Gaviano M., Lera D., Kvasov D.E., Sergeyev Y.D.} Software for generation of classes of test functions with known local and global minima for global optimization.
+ACM Transactions on Mathematical Software.
+– 2003.
+– Vol.
+29.
+– P.
+469-480.}
 
-\elitem{rfa:enlit:mrow}{%
-Multirow~--- Create Tabular Cells Spanning Multiple Rows.
-\url{https://www.ctan.org/pkg/multirow}.
-Cited November~12, 2021.}
+\elitem{rfa:rulit:lit8}{
+\textit{Sergeev Ya.D., Kvasov D.E.} %Диагональные методы глобальной оптимизации.  М.: Физматлит, 2008. – 352 c.
+}
+
+\elitem{rfa:rulit:lit9}{
+\textit{Gablonsky J.M., Kelley C.T.} A Locally-Biased Form of the DIRECT Algorithm. Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.}
+
+\elitem{rfa:rulit:lit10}{
+\textit{Paulavicius R., Zilinskas J, Grothey A.} Parallel branch and bound for global optimization with combination of Lipschitz bounds.
+Optimization Methods \& Software,  2011.
+– Vol.
+26, No.
+3.
+P.
+487–498.}
+
+\elitem{rfa:rulit:lit11}{
+\textit{Boreskov A.A., Kharlamov A.A., Markovsky N.D., Mikushin D.N., Mortikov E.V., Myltsev A.A., Sakharnykh N.A., Frolov V.A.} %Параллельные вычисления на GPU.
+Архитектура и программная модель CUDA.
+М.:  Издательство Московского университета, 2015.
+336 с.
+}
+
+\elitem{rfa:rulit:lit12}{
+\textit{Tanenbaum A., Austin T.} Structured computer organization }
+
+\elitem{rfa:rulit:lit13}{
+\textit{Komolov D.A., Myalk R.A., Zobenko A.A., Filippov A.S.} %Системы автоматизированного проектирования фирмы Altera MAX+Plus II и Quartus II. Издательство РадиоСофт, 2002. 355 c.
+}
+
+\elitem{rfa:rulit:lit14}{
+\textit{Gergel V.P., Barkalov K.A., Sysoyev A.V.} A novel supercomputer software system for solving time-consuming global optimization problems.
+Numerical Algebra, Control and Optimization, 2018.
+Vol.
+8, No.
+1.
+P.
+47-62.}
+
+\elitem{rfa:rulit:lit15}{
+\textit{Sysoyev, A., Barkalov, K., Sovrasov, V., Lebedev, I., Gergel, V.} AGS NLP solver. \url{ https://github.com/sovrasov/ags_nlp_solver}}
+
+\elitem{rfa:rulit:lit16}{
+\textit{Sergeev, Ya.D., Kvasov D.E.} %Диагональные методы глобальной оптимизации. М.: Физматлит, 2008. 352 c.
+}
 

### 2022-10-11 2022/PAVT Special Issue/mathematics-1918483 -for proof.tex

--- 2022-09-22 2022/PAVT Special Issue/template.tex
@@
 %=================================================================
-\documentclass[mathematics,article,submit,pdftex,moreauthors]{Definitions/mdpi} 
+\documentclass[mathematics,article,accept,pdftex,moreauthors]{Definitions/mdpi} 
 % For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
@@
 \copyrightyear{2022}
-%\externaleditor{Academic Editor: Firstname Lastname}
-\datereceived{} 
+\externaleditor{Academic Editor: {Leonid B. Sokolinsky, Mikhail Zymbler}}
+
+\datereceived{31 August 2022} 
 %\daterevised{} % Only for the journal Acoustics
-\dateaccepted{} 
+\dateaccepted{4 October 2022 } 
 \datepublished{} 
@@
 % Full title of the paper (Capitalized)
-\Title{On Solving the Problem of Finding Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction Using Parallel Global Search Algorithm}
-% Дежурное название: Использование ...алгоритма для определения кинетических параметров процесса каталитической изомеризации пентан-гексановой фракции
-% Еще один вариант: О решении задачи определения Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction с помощью Parallel Global Search Algorithm
+\Title{On Solving the Problem of Finding Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction Using~a Parallel Global Search Algorithm}
+
 % На английском: On Solving the Problem of Finding Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction Using Parallel Global Search Algorithm
@@
 % MDPI internal command: Title for citation in the left column
-\TitleCitation{On Solving the Problem of Finding Kinetic Parameters}
+\TitleCitation{On Solving the Problem of Finding Kinetic Parameters of Catalytic Isomerization of the Pentane-Hexane Fraction Using a Parallel Global Search Algorithm}
 
@@
 % Authors, for the paper (add full first names)
-\Author{Konstantin Barkalov $^{1}$\orcidA{}, Irek Gubaydullin $^{2,3}$, Evgeny Kozinov$^{1}$\orcidB{}, Ilya Lebedev $^{1}$\orcidC{}, Roza Faskhutdinova\orcidD{} $^{2,3}$, Azamat Faskhutdinov$^{2}$ and Leniza Enikeeva $^{3,4,}$*\orcidE{}}
+\Author{\hl{Konstantin Barkalov} %MDPI: 1.
+Please carefully check the accuracy of names and affiliations.
+2.
+Please do not delete our comments.
+3.
+Please revise all questions that we proposed.
+And reply to them after the corresponding comment.
+Such as: “Author reply: It should be italic”; “Author reply: I confirm”; “Author reply: I have checked and revised all.” etc..
+ $^{1}$\orcidA{}, Irek Gubaydullin $^{2,3}$, Evgeny Kozinov $^{1}$\orcidB{}, Ilya Lebedev $^{1}$\orcidC{}, Roza Faskhutdinova\orcidD{} $^{2,3}$, Azamat Faskhutdinov $^{2}$ and Leniza Enikeeva $^{3,4,}$*\orcidE{}}
 
@@
 \address{%
-$^{1}$ \quad Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia; konstantin.barkalov@itmm.unn.ru\\
-$^{2}$ \quad Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia; irekmars@mail.ru \\
-$^{3}$ \quad Ufa State Petroleum Technological University, Ufa, Russia; roza\_fask@mail.ru\\
-$^{4}$ \quad Novosibirsk State University, Novosibirsk, Russia; leniza.enikeeva@yandex.ru
+	$^{1}$ \quad \hl{Lobachevsky} %MDPI: Please add the department/school/faculty/campus in affs. 1, 3 and 4.
+ State University of Nizhny Novgorod, \hl{Nizhny Novgorod,} %MDPI: Please add the postal code in affs. 1--4 (or ZIP code in the U.S.). If the postal code is not available, please provide the P.O. Box.
+ Russia; konstantin.barkalov@itmm.unn.ru (K.B.); evgeny.kozinov@itmm.unn.ru (E.K.); ilya.lebedev@itmm.unn.ru~(I.L.)\\
+	$^{2}$ \quad \hl{Institute} %MDPI: If the information provided presents more than one address, please separate the addresses into different affiliations.
+ of Petrochemistry and Catalysis---Subdivision of the Ufa Federal Research Centre of RAS, \mbox{\hl{Ufa}, Russia;} irekmars@mail.ru (I.G.); roza\_fask@mail.ru (R.F.); azamat\_d1989@mail.ru (A.F.)\\
+	$^{3}$ \quad Ufa State Petroleum Technological \hl{University}, \hl{Ufa}, Russia\\
+	$^{4}$ \quad Novosibirsk State \hl{University}, \hl{Novosibirsk}, Russia
 }
@@
 % Contact information of the corresponding author
-\corres{Correspondence: leniza.enikeeva@yandex.ru (E.L.); konstantin.barkalov@itmm.unn.ru (B.K.) }
+\corres{Correspondence: leniza.enikeeva@yandex.ru}
 
@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{The article is devoted to the problem of developing a kinetic model of a complex chemical reaction using a parallel optimization method.
-The design of the kinetic model consists in finding the kinetic parameters of the reaction, which cannot be calculated analytically, and since the chemical reaction involves many stages, the optimization problem is multi-extreme.
+\abstract{This article is devoted to the problem of developing a kinetic model of a complex chemical reaction using a parallel optimization method.
+The design of the kinetic model consists of finding the kinetic parameters of the reaction, which cannot be calculated analytically, and since the chemical reaction involves many stages, the optimization problem is {multiextremal}.
 As a chemical reaction, the process of catalytic isomerization of the pentane-hexane fraction is considered, which is now important due to the switch of the oil refining industry to the production of gasoline corresponding to the Euro-5 standard.
-To solve the optimization problem, the authors have developed parallel global search algorithm and a program based on Lipschitz optimization.
-The kinetic parameters found made it possible to develop a mathematical model of the process, which gives good agreement with industrial data.}
+{On the basis of known industrial data on the concentrations of reaction components and the temperature at the outlet of the third reactor, the activation energies and pre-exponential factors of each reaction stage were calculated.} To solve the optimization problem, the authors developed a parallel global search algorithm and a program based on Lipschitz optimization.
+The kinetic parameters found made it possible to develop a mathematical model of the process, which is in good agreement with industrial data.
+{The developed mathematical model in future works will make it possible to study the dynamics of the gas--liquid flow in the reactor unit, taking into account diffusion and heat exchange processes through the catalyst layer.}}
 
 % Keywords
-\keyword{kinetic model; inverse problems of chemical kinetics; mathematical modeling; catalytic isomerisation of pentane-hexane fraction; Lipschitz optimization; black-box functions; parallel computing}
+\keyword{inverse problems of chemical kinetics; mathematical modeling; catalytic isomerisation; Lipschitz optimization; black-box functions; parallel computing}
 
@@
 %\PACS{J0101}
-%\MSC{}
+\MSC{90C26; 65Y05}
 %\JEL{}
@@
 
-Due to the transition of the domestic oil refining industry to the production of gasoline conforming to the Euro-5 standard, an urgent task is to reduce the content of aromatic hydrocarbons and, in particular, benzene in motor fuel while maintaining the octane value.
-In connection with the transition of the oil refining industry to the production of motor gasoline conforming to Euro-5 and Euro-6 standards, the reduction of aromatic hydrocarbons in motor fuel and, mainly, benzene becomes an urgent task.
-At the same time, removing aromatic hydrocarbons, which are high-octane components, it is necessary to preserve the octane number of gasoline.
-It is known that hydrocarbons of branched structure have a higher octane number than hydrocarbons of linear structure.
-For example, in $n$-pentane, the octane number according to the research method is 62 points, and in $iso$-pentane it is 93 points.
+	Due to the transition of the domestic oil refining industry to the production of gasoline conforming to the Euro-5 standard, reducing the content of aromatic hydrocarbons is an urgent task, in particular benzene in motor fuel, while maintaining the octane value.
+In connection with the transition of the oil refining industry to the production of motor gasoline conforming to Euro-5 and Euro-6 standards, the reduction of aromatic hydrocarbons in motor fuel and, mainly, benzene, becomes an urgent task.
+At the same time, when removing aromatic hydrocarbons, which are high-octane components, it is necessary to preserve the octane number of gasoline.
+It is known that hydrocarbons with a branched structure have a higher octane number than hydrocarbons with a linear structure.
+For example, in $n$-pentane, the octane number according to the research method is 62 points and in $iso$-pentane it is 93 points.
 On an industrial scale, catalytic isomerization units of light paraffins have been introduced to produce hydrocarbons of an isomeric structure.
@@
 The main advantage of this process is the production of high-octane gasoline, while aromatic hydrocarbons are practically not formed.
-Straight-run gasoline fractions with boiling limits of 62-70 $^0$C containing mainly $n$-pentane and $n$-hexane, pentane-hexane fractions from gas condensate or from gas fractionation plants, as well as catalytic reforming raffinates are used as feedstock for catalytic isomerization.
+Straight-run gasoline fractions with boiling limits of 62--70 °C containing mainly $n$-pentane and $n$-hexane, pentane-hexane fractions from gas condensate or from gas fractionation plants, as well as catalytic reforming raffinates, are used as feedstock for catalytic isomerization.
 Modern bifunctional catalysts are used to carry out the process.
-In order to prevent coke deposition, the catalytic isomerization process is carried out in a hydrogen-containing gas environment under pressure.
+In order to prevent coke deposition, the catalytic isomerization process is carried out in a hydrogen-containing gas environment under pressure \cite{Hartmut, Fahim, Meyers}.
 
 	Computer information technologies are widely used in the modeling of production processes and today it is difficult to imagine research activities without the use of modern means of constructing and using models.
-Moreover, special attention is currently being paid to the development of a detailed kinetic model of complex industrial processes.
-Based on these models, it is possible to significantly increase the efficiency of the process, while minimally spending energy and material resources.
+Moreover, special attention is currently being paid to the development of a detailed kinetic model of complex industrial processes \cite{DAO2022172, SAJID20207430, LI2021175, HOU201759}.
+Based on these models, it is possible to significantly increase the efficiency of the process while utilizing minimal energy and material resources \cite{uskovfibrous, en13133393}.
 The main advantage of such methods is the absence of additional costs for the construction of pilot plants.
-The developed model should describe the process within a wide range of operating conditions, have high flexibility, allowing for changes, calculations of the composition of the reaction mixture, etc.
+The developed model should describe the process within a wide range of operating conditions, have high flexibility and allow for changes and calculations of the composition of the reaction mixture, etc.
 
-In the article \cite{enik2021} we have already reviewed this process, but in order to simplify the model, the scheme of chemical transformations, in particular hydrocracking reactions, have been revised.
-Thus, the most probable hydrocracking reactions based on the carbonium-ion mechanism were formed.
+	{The reactor unit of the catalytic isomerization unit of the pentane-hexane fraction is the object of research in this work.
+The production process was carried out in a continuous way.
+The reactor unit consisted of three consecutive reactors.
+		The raw material was a light gasoline fraction after hydrotreating.
+The catalyst of the process was a bifunctional catalyst of the SI-2 brand.
+The total mass of the catalyst in the reactor unit was 27 tons (9 tons in each reactor).
+The process was carried out at a constant pressure of 3.2 MPa.
+		For calculations, data for three days of operation of an industrial installation were taken.
+		Based on chromatographic analyses and mass flow rates for each day, the component compositions of the reaction mixture were formed both at the entrance to the reactor unit and at the exit from the reactor~unit.}
+
+	In \cite{enik2021} we already reviewed this process, but in order to simplify the model, the scheme of chemical transformations, in particular hydrocracking reactions, have been revised.
+Thus, the most probable hydrocracking reactions, based on the carbonium-ion mechanism, were formed.
 The new transformation scheme implies the recalculation of the kinetic parameters of the reaction, which form a mathematical model of the process.
 
-In the article \cite{enik2021}, a genetic algorithm was used to calculate kinetic parameters, and the calculation time of kinetic parameters was about 10 hours.
+	In \cite{enik2021}, a genetic algorithm was used to calculate kinetic parameters, and the calculation time of kinetic parameters was about 10 h.
 Therefore, in order to develop a mathematical model of a new transformation scheme, the question of choosing a more effective optimization method arises.
 
-In the complex problems of chemical kinetics, kinetic constants of reactions cannot be found analytically, they can only be found using numerical methods (see, for example, \cite{Zaynullin2020,Enikeeva2020,Koledina2019}).
-In this case, the quality criterion (objective function) is set in the form of a certain algorithm, which includes the stage of numerical modeling, which, in turn, makes it compute-intensive.
+	In the complex problems of chemical kinetics, kinetic constants of reactions cannot be found analytically; they can only be found using numerical methods (see, for example,~\cite{Enikeeva2020}).
+In this case, the quality criterion (objective function) is set in the form of a certain algorithm, which includes the stage of numerical modeling, which, in turn, makes it computationally intensive.
 Furthermore, in the inverse problems of chemical kinetics, the objective function can be substantially multiextremal, i.e., having many local extremes along the global one.
+	{Gradient-based optimization methods will not provide a good result for such problems.
+On the one hand, such methods converge only to a local solution.
+On the other hand, in the considered problem, the gradient of the objective function is not known analytically and its numerical estimate will be a time-consuming operation.}
 
@@
 Metaheuristic algorithms are usually based on simulations of processes occurring in the wild.
-Examples of metaheuristic algorithms are simulated annealing, evolutionary computing, etc. (see, for example, \cite{Battiti2009,Eiben2015}).
-These algorithms are included in many software libraries for solving global optimization problems (for example, MATLAB Global Optimization Toolbox, Scipy.Optimize, etc.).
+	Examples of metaheuristic algorithms are simulated annealing and evolutionary computing, etc. (see, for example, \cite{Battiti2009,Eiben2015}).
+	These algorithms are included in many software libraries for solving global optimization problems (for example, MATLAB Global Optimization Toolbox, \hl{Scipy.Optimize,} %MDPI: Please check if there should be a space after ``Scipy.''.
+ etc.).
 	Due to the operation speed and low requirements for computing resources (in particular, memory), metaheuristic algorithms have become widespread for solving practical problems.
-However, the solution to the problem found by such an algorithm is, in general, local and can be located far from the global optimum \cite{Kvasov2018}.
+However, the solution to the problem found by such an algorithm is, in general, local, and can be located far from the global optimum \cite{Kvasov2018}.
 
@@
 It is assumed that the values of the objective function are determined as a result of solving some other task and require a costly computational experiment.
+	{On many benchmarks, Lipschitz global optimization methods outperform metaheuristic algorithms in terms of the number of correctly solved problems~\cite{Sergeyev2018,Sovrasov2019}.}
 
-This paper reflects the results of applying parallel methods of Lipschitz optimization developed in \cite{Barkalov2016,Strongin2018,Sysoyev2017}, to the solution of inverse problems of chemical kinetics.
+	This paper reflects the results of applying parallel methods of Lipschitz optimization developed in \cite{Barkalov2016,Strongin2018,Sysoyev2017} to the solution of inverse problems of chemical kinetics.
 	The main part of the paper is structured as follows.
@@
 Section \ref{sec_numexper} outlines the issues related to the technical aspects of the implementation of the parallel algorithm.
-The physico-chemical interpretation of the results of numerical calculations using the developed parallel algorithm is presented in the Section \ref{sec_discussion}.
+The physico-chemical interpretation of the results of numerical calculations using the developed parallel algorithm is presented in Section \ref{sec_discussion}.
 	%The last section contains the conclusion and plans for future work.
 
-\section{Mathematical model}\label{sec_matmodel}
+	\section{Mathematical Model}\label{sec_matmodel}
 
-Let's consider the mechanism of chemical transformations of the isomerization of the pentane-hexane fraction.
-The reaction consists of 48 reaction stages, which are presented in the table \ref{tab_chemistry}.
+	Let us consider the mechanism of chemical transformations of the isomerization of the pentane-hexane fraction.
+The reaction consists of 48 reaction stages, which are presented in Table \ref{tab_chemistry}.
 The following abbreviations were adopted: 2-MP is 2-methylpentane; 3-MP is 3-methylpentane; 2,2-DMB is 2,2-dimethylbutane; 2,3-DMB is 2,3-dimethylbutane; CH* is cyclohexane; MCP is methyl cyclopentane; B is benzene; CP is cyclopentane.
+
+	The mathematical model of the chemical process is a non-linear system of differential equations \cite{enik2021,logan}:
+	\begin{linenomath}
+		\begin{equation}\label{eq_concentrations}
+			\dfrac{dx_i}{d\tau} = \sum_{j=1}^J \nu_{ij} w_j, i = 1,\dots,I,
+		\end{equation}
+		\begin{equation}\label{eq_rate}
+			w_{j} = k_j \prod_{i=1}^I \left( \dfrac{x_i}{F} \right),
+		\end{equation}
+		\begin{equation}\label{eq_constant}
+			k_j = k^0_j \exp \left( -\dfrac{E_j}{RT} \right),
+		\end{equation}
+	\end{linenomath}
+	with initial values at $\tau = 0$, $x_i(0) = x_i^0$, where $\nu_{ij}$ are stoichiometric coefficients of stages of the chemical reaction; $J$ is the number of stages ($J = 48$), $x_i$ is the molar flow rate of the $i$-th component of reaction, kmol/h; $I$ is the number of components ($I = 17$); $w_j$ is the rate of the $j$-th stage of reaction, kmol/(h$\cdot$kg cat.), $E_j$ is the activation energy of the reaction stage, J/mol; $R$ is the universal gas constant ($R$ = 8.31 J/(mol$\cdot$K), $T$ is the temperature, K; $k_j$ are pre-exponential factors, kmol/(h$\cdot$kg cat.); $\tau$ is conditioned contact time, (kg cat.); $F$ is the total molar flow rate, kmol/h.
+
+	The mathematical model involves parameters such as the raw material flow rate $F$ in the reactor and temperature $T$, but these two parameters also depend on $\tau$; therefore it is necessary to write equations on them because the technological process is non-isothermal:
+	\begin{linenomath}
+		\begin{equation}\label{eq_temper}
+			\dfrac{dT}{d\tau} = \dfrac{\sum_{i=1}^I\left( \dfrac{dx_i}{d\tau} \cdot \Delta H^i_{f,T} \right) }{\sum_i^J \left(x_i C^T_{p_i} \right) },
+		\end{equation}
+		\begin{equation}\label{eq_flow}
+			\dfrac{dF}{d\tau} = \sum_{i=1}^I  \dfrac{x_i}{d\tau},
+		\end{equation}
+	\end{linenomath}
+	with initial condition: at $\tau = 0$, $T(0) = T^0$, where $T$ is the temperature, K, $\Delta H^i_{f,T}$ is the enthalpy of the $i$-th component formation at temperature $T$, J/mol, $C^T_{p_i}$ is the specific heat capacity of the $i$-th component at temperature $T$, (J/mol$\cdot$K), $F$ is the mole discharge of the flow, kmol/h.
 
@@
 			\toprule
- \textbf{Reaction stages} & \textbf{Reaction stages} \\
+			\textbf{Reaction Stages} & \textbf{Reaction Stages} \\
 			\midrule
-1) $n$-C$_5$H$_{12}$ $\rightarrow$ $iso$-C$_5$H$_{12}$ & 25) $n$-C$_6$H$_{14}$ $\rightarrow$ CH* + H$_2$\\
-2) $iso$-C$_5$H$_{12}$ $\rightarrow$ $n$-C$_5$H$_{12}$ & 26) CH* + H$_2$ $\rightarrow$ $n$-C$_6$H$_{14}$ \\
-3) $n$-C$_6$H$_{14}$ $\rightarrow$ 2-MP & 27) 2-MP $\rightarrow$ MCP+ H$_2$ \\
-4) 2-MP $\rightarrow$ $n$-C$_6$H$_{14}$ & 28) MCP + H$_2$ $\rightarrow$ 2-MP  \\
-5) $n$-C$_6$H$_{14}$ $\rightarrow$ 3-MP & 29) 3-MP $\rightarrow$ MCP + H$_2$ \\
-6) 3-MP $\rightarrow$ $n$-C$_6$H$_{14}$  & 30) MCP + H$_2$ $\rightarrow$ 3-MP \\
-7)  2-MP $\rightarrow$ 3-MP & 31) 2,2-DMB $\rightarrow$ MCP + H$_2$ \\
-8) 3-MP $\rightarrow$ 2-MP  & 32) MCP + H$_2$ $\rightarrow$ 2,2-DMB \\
-9) 2-MP $\rightarrow$ 2,2-DMB  & 33) 2,3-DMB $\rightarrow$ MCP + H$_2$ \\
-10) 2,2-DMB $\rightarrow$ 2-MP  & 34) MCP + H$_2$ $\rightarrow$ 2,3-DMB \\
-11) 2-MP $\rightarrow$ 2,3-DMB  & 35) CP + H$_2$ $\rightarrow$ $n$-C$_5$H$_{12}$ \\
-12) 2,3-DMB $\rightarrow$ 2-MP  & 36) $n$-C$_5$H$_{12}$ $\rightarrow$ CP + H$_2$ \\
-13) 3-MP $\rightarrow$ 2,2-DMB  & 37) $n$-C$_4$H$_{10}$ $\rightarrow$ $iso$-C$_4$H$_{10}$ \\
-14) 2,2-DMB $\rightarrow$ 3-MP  & 38) $iso$-C$_4$H$_{10}$ $\rightarrow$ $n$-C$_4$H$_{10}$ \\
-15) 3-MP $\rightarrow$ 2,3-DMB  & 39) $n$-C$_5$H$_{12}$ + H$_2$ $\rightarrow$ C$_3$H$_8$ + C$_2$H$_6$ \\
-16) 2,3-DMB $\rightarrow$ 3-MP  & 40) $n$-C$_5$H$_{12}$ + H$_2$ $\rightarrow$ $n$-C$_4$H$_{10}$ + CH$_4$ \\
-17) 2,2-DMB $\rightarrow$ 2,3-DMB  & 41) $iso$-C$_5$H$_{12}$ + H$_2$ $\rightarrow$ $iso$-C$_4$H$_{10}$ + CH$_4$ \\
-18) 2,3-DMB $\rightarrow$ 2,2-DMB & 42) $n$-C$_6$H$_{14}$ + H$_2$ $\rightarrow$ 2C$_3$H$_8$ \\
-19) B + 3H$_2$ $\rightarrow$ CH*  & 43) $n$-C$_6$H$_{14}$ + H$_2$ $\rightarrow$ $n$-C$_5$H$_{12}$ + CH$_4$ \\
-20) CH* $\rightarrow$ B + 3H$_2$ & 44) $n$-C$_6$H$_{14}$ + H$_2$ $\rightarrow$ $n$-C$_4$H$_{10}$ + C$_2$H$_6$ \\
-21) B + 3H$_2$ $\rightarrow$ MCP  & 45) 2-MP + H$_2$ $\rightarrow$ $iso$-C$4$H$_{10}$ + C$_2$H$_6$ \\
-22) MCP $\rightarrow$ B + 3H$_2$  & 46) 3-MP + H$_2$ $\rightarrow$ $iso$-C$_5$H$_{12}$ + CH$_4$ \\
-23) CH* $\rightarrow$ MCP & 47) 2,2-DMB + H$_2$ $\rightarrow$ $iso$-C$_5$H$_{12}$ + CH$_4$ \\
-24) MCP $\rightarrow$ CH*  & 48) 2,3-DMB + H$_2$ $\rightarrow$ $iso$-C$_5$H$_{12}$ + CH$_4$ \\
+			(1) $n$-C$_5$H$_{12}$$\rightarrow$$iso$-C$_5$H$_{12}$ & (25) $n$-C$_6$H$_{14}$$\rightarrow$CH* + H$_2$\\
+			(2) $iso$-C$_5$H$_{12}$$\rightarrow$$n$-C$_5$H$_{12}$ & (26) CH* + H$_2$$\rightarrow$$n$-C$_6$H$_{14}$ \\
+			(3) $n$-C$_6$H$_{14}$$\rightarrow$2-MP & (27) 2-MP$\rightarrow$MCP+ H$_2$ \\
+			(4) 2-MP$\rightarrow$$n$-C$_6$H$_{14}$ & (28) MCP + H$_2$$\rightarrow$2-MP  \\
+			(5) $n$-C$_6$H$_{14}$$\rightarrow$3-MP & (29) 3-MP$\rightarrow$MCP + H$_2$ \\
+			(6) 3-MP$\rightarrow$$n$-C$_6$H$_{14}$  & (30) MCP + H$_2$$\rightarrow$3-MP \\
+			(7)  2-MP$\rightarrow$3-MP & (31) 2,2-DMB$\rightarrow$MCP + H$_2$ \\
+			(8) 3-MP$\rightarrow$2-MP  & (32) MCP + H$_2$$\rightarrow$2,2-DMB \\
+			(9) 2-MP$\rightarrow$2,2-DMB  & (33) 2,3-DMB$\rightarrow$MCP + H$_2$ \\
+			(10) 2,2-DMB$\rightarrow$2-MP  & (34) MCP + H$_2$$\rightarrow$2,3-DMB \\
+			(11) 2-MP$\rightarrow$2,3-DMB  & (35) CP + H$_2$$\rightarrow$$n$-C$_5$H$_{12}$ \\
+			(12) 2,3-DMB$\rightarrow$2-MP  & (36) $n$-C$_5$H$_{12}$$\rightarrow$CP + H$_2$ \\
+			(13) 3-MP$\rightarrow$2,2-DMB  & (37) $n$-C$_4$H$_{10}$$\rightarrow$$iso$-C$_4$H$_{10}$ \\
+			(14) 2,2-DMB$\rightarrow$3-MP  & (38) $iso$-C$_4$H$_{10}$$\rightarrow$$n$-C$_4$H$_{10}$ \\
+			(15) 3-MP$\rightarrow$2,3-DMB  & (39) $n$-C$_5$H$_{12}$ + H$_2$$\rightarrow$C$_3$H$_8$ + C$_2$H$_6$ \\
+			(16) 2,3-DMB$\rightarrow$3-MP  & (40) $n$-C$_5$H$_{12}$ + H$_2$$\rightarrow$$n$-C$_4$H$_{10}$ + CH$_4$ \\
+			(17) 2,2-DMB$\rightarrow$2,3-DMB  & (41) $iso$-C$_5$H$_{12}$ + H$_2$$\rightarrow$$iso$-C$_4$H$_{10}$ + CH$_4$ \\
+			(18) 2,3-DMB$\rightarrow$2,2-DMB & (42) $n$-C$_6$H$_{14}$ + H$_2$$\rightarrow$2C$_3$H$_8$ \\
+			(19) B + 3H$_2$$\rightarrow$CH*  & (43) $n$-C$_6$H$_{14}$ + H$_2$$\rightarrow$$n$-C$_5$H$_{12}$ + CH$_4$ \\
+			(20) CH*$\rightarrow$B + 3H$_2$ & (44) $n$-C$_6$H$_{14}$ + H$_2$$\rightarrow$$n$-C$_4$H$_{10}$ + C$_2$H$_6$ \\
+			(21) B + 3H$_2$$\rightarrow$MCP  & (45) 2-MP + H$_2$$\rightarrow$$iso$-C$4$H$_{10}$ + C$_2$H$_6$ \\
+			(22) MCP$\rightarrow$B + 3H$_2$  & (46) 3-MP + H$_2$$\rightarrow$$iso$-C$_5$H$_{12}$ + CH$_4$ \\
+			(23) CH*$\rightarrow$MCP & (47) 2,2-DMB + H$_2$$\rightarrow$$iso$-C$_5$H$_{12}$ + CH$_4$ \\
+			(24) MCP$\rightarrow$CH* & (48) 2,3-DMB + H$_2$$\rightarrow$$iso$-C$_5$H$_{12}$ + CH$_4$ \\
 			\bottomrule
@@
 
-Mathematical model of the chemical process is a non-linear system of differential equations \cite{Koledina2016KineticsAM,enik2021,logan}:
+	Thus, when solving the inverse kinetic problem, the direct problem is repeatedly solved, which includes the numerical integration of Equations \eqref{eq_concentrations}, \eqref{eq_temper} and \eqref{eq_flow}.
+After finding the numerical values of the molar flow rate $x_i$, temperature $T$ and total molar flow rate $F$, these values are compared with the corresponding industrial data obtained during the three days of operation of the catalytic isomerization unit \cite{enik2021}, which is a minimization~task:
+
 	\begin{linenomath}
-\begin{equation}\label{eq_concentrations}
-\dfrac{dx_i}{d\tau} = \sum_{j=1}^J \nu_{ij} w_j, i = 1,...,I,
-\end{equation}
-\begin{equation}\label{eq_rate}
-w_{j} = k_j \prod_{i=1}^I \left( \dfrac{x_i}{F} \right),
-\end{equation}
-\begin{equation}\label{eq_constant}
-k_j = k^0_j \exp \left( -\dfrac{E_j}{RT} \right)
+		\begin{equation}\label{FF}
+			FF = \dfrac{\sum_{i=1}^I \left| \dfrac{y_i^{exp} - y_i^{calc}}{y_i^{exp}} \right|}{I}\rightarrow\min,
 		\end{equation}
 	\end{linenomath}
-with initial values: at $\tau = 0$, $x_i(0) = x_i^0$, where $\nu_{ij}$ are stoichiometric coefficients of stages of chemical reaction; $J$ is number of stages ($J = 48$), $x_i$ is molar flow rate of the $i$-th component of reaction, kmol/h; $I$ is the number of components ($I = 17$); $w_j$ is the rate of the $j$-th stage of reaction, kmol/(h$\cdot$kg cat.), $E_j$ is activation energy of reaction stage, J/mol; $R$ is the universal gas constant ($R$=8,31 J/(mol$\cdot$K), $T$ is temperature, K; $k_j$ are pre-exponentional factors, kmol/(h$\cdot$kg cat.); $\tau$ is conditioned contact time, (kg cat.); $F$ is the total molar flow rate, kmol/h.
-
-The mathematical model involves such parameters as the raw material flow rate $F$ in the reactor and temperature $T$, but these two parameters also depend on $\tau$, therefore it is necessary to write equations on them, because technological process is non-isothermal:
-\begin{linenomath}
-\begin{equation}\label{eq_temper}
-\dfrac{dT}{d\tau} = \dfrac{\sum_{i=1}^I\left( \dfrac{dx_i}{d\tau} \cdot \Delta H^i_{f,T} \right) }{\sum_i^J \left(x_i C^T_{p_i} \right) },
-\end{equation}\label{eq_flow}
-\begin{equation}
-\dfrac{dF}{d\tau} = \sum_{i=1}^I  \dfrac{x_i}{d\tau}
-\end{equation}
-\end{linenomath}
-with initial condition: at $\tau = 0$, $T(0) = T^0$, where $T$ is temperature, K, $\Delta H^i_{f,T}$ is enthalpy of the $i$-th component formation at temperature $T$, J/mol, $C^T_{p_i}$ is specific heat capacity of the $i$-th component at temperature $T$, (J/mol$\cdot$K), $F$ is mole discharge of the flow, kmol/h.
-
-Thus, when solving the inverse kinetic problem, the direct problem is repeatedly solved, which includes the numerical integration of equations \eqref{eq_concentrations}, \eqref{eq_temper} and \eqref{eq_flow}.
-After finding the numerical values of the molar flow rate $x_i$, temperature $T$ and total molar flow rate $F$, these values are compared with the corresponding industrial data obtained during the three days of operation of the catalytic isomerization unit \cite{enik2021}, which is a minimization task:
-
-\begin{linenomath}
-\begin{equation}
-FF = \dfrac{\sum_{i=1}^I \left| \dfrac{y_i^{exp} - y_i^{calc}}{y_i^{exp}} \right|}{I} \rightarrow \min,
-\end{equation}
-\end{linenomath}
-
-where I is the number of observed parameters (I=19, of which 17 parameters are $x_i$, one parameter is $T$ and one parameter is $F$),  $y_{i}^{exp}$ are industrual data, $y_{i}^{calc}$ are numerical values of concentrations of reaction substances obtained by solving a direct kinetic problem.
-As a result of solving the inverse problem, the kinetic parameters $k^0_j$ and $E_j$ included in the Arrhenius equation \eqref{eq_rate} will be obtained.
-Since the number of reaction stages is 48, the kinetic model will contain 96 kinetic parameters (activation energy $E_j$ and pre-exponential multiplier $k^0_j$ for each stage), however, the authors decided to fix some of the parameters as already known for chemical reasons, so the number of optimized parameters was reduced to 24.
+	where \emph{I} is the number of observed parameters (\emph{I} = 19, of which 17 parameters are $x_i$, one parameter is $T$ and one parameter is $F$),  $y_{i}^{exp}$ are the industrial data, $y_{i}^{calc}$ are the numerical values of concentrations of reaction substances obtained by solving a direct kinetic problem.
+As a result of solving the inverse problem, the kinetic parameters $k^0_j$ and $E_j$ included in the Arrhenius Equation \eqref{eq_rate} will be obtained.
+Since the number of reaction stages is 48, the kinetic model will contain 96 kinetic parameters (activation energy $E_j$ and pre-exponential multiplier $k^0_j$ for each stage); however, the authors decided to fix some of the parameters as already known for chemical reasons, so the number of optimized parameters was reduced to 24.
 
@@
 
-\subsection{Statement of the global optimization problem}
+	\subsection{Statement of the Global Optimization Problem}
 
 	As stated above, from a mathematical point of view, the inverse problem of chemical kinetics can be considered a Lipschitz global optimization problem.
+	{This fact has has the following justification.
+		We solve the ODE system (\ref{eq_concentrations})--(\ref{eq_flow}) to compute the value of the objective function (\ref{FF}).
+The right-hand side of the ODE system is continuous functions with bounded derivatives.
+Thus, the system solution will also be continuous and bounded at the bounded period of time.
+Therefore, the objective function (\ref{FF}) will satisfy the Lipschitz condition, but the Lipschitz constant will be a priori unknown.}
+
 	In general, the problem of the specified class can be formulated as follows:
@@
 	\end{gather}
-where $a,b \in R^N$ are boundaries of the search domain, and the objective function $\phi(y)$ satisfies the Lipschitz condition
+	where $a,b \in R^N$ are boundaries of the search domain, and the objective function $\phi(y)$ satisfies the Lipschitz condition,
 	\begin{equation}\label{Lip}
@@
 	The criterion $\phi(y)$ is assumed to be multiextremal and is given in the form of some computational algorithm. 
+	{This assumption is typical for many applied optimization problems (see, e.g., \cite{Kvasov2013,Kalyulin2017}).}
 	The input of this algorithm is supplied with a parameter vector, and the output is the corresponding function value.
@@
 
-One of the main difficulties in solving Lipschitz global optimization problems is the so-called ``curse of dimensionality'' --- the exponential increase in the number of trials required to solve a problem with a given accuracy, accompanied by the increasing dimensionality of the problem.
-It is possible to smooth this effect while maintaining the accuracy of the solution by constructing significantly uneven coverages of the search domain with trial points.
+	One of the main difficulties in solving Lipschitz global optimization problems is the so-called ``curse of dimensionality''---the exponential increase in the number of trials required to solve a problem with a given accuracy, accompanied by the increasing dimensionality of the problem.
+It is possible to smooth this effect while maintaining the accuracy of the solution by constructing significantly uneven coverages of the search domain with trial~points.
 
-Known examples of such methods are DIvide RECTangles (DIRECT) algorithm \cite{Jones1993}, the simplicial partitions algorithm \cite{PaulaviciusZilinskas2014}, the diagonal partitions algorithm \cite{Sergeyev2017}.
+	Known examples of such methods are the DIvide RECTangles (DIRECT) algorithm \cite{Jones1993}, the simplicial partitions algorithm \cite{PaulaviciusZilinskas2014}, the diagonal partitions algorithm \cite{Sergeyev2017}.
 These techniques are based on adaptive partitioning of the search domain into a finite number of subdomains at each iteration.
-Then the need for further subdivision for a more detailed study of these subdomains is estimated (using the accumulated information about the objective function). 
+	Then, the need for further subdivision for a more detailed study of these subdomains is estimated (using the accumulated information about the objective function). 
 	A fundamentally different approach to solving a multidimensional problem (\ref{problemN}) is its reduction to one or more one-dimensional problems followed by the use of efficient methods of one-dimensional optimization.
-Such reduction may be accomplished, for example, by a recursive optimization scheme \cite{Grishagin2018} or by Peano curves \cite{Barkalov2018}.
+Such a reduction may be accomplished, for example, by a recursive optimization scheme \cite{Grishagin2018} or by Peano curves \cite{Sergeyev2013}.
 The latter approach is used in this work.
 
-Using a continuous single-valued correspondence (Peano curve) $y(x)$ of the real axis segment $[0,1]$ onto hyperinterval $D$ from (\ref{D}), a multidimensional problem (\ref{problemN}) can be reduced to a one-dimensional one
+	Using a continuous single-valued correspondence (Peano curve) $y(x)$ of the real axis segment $[0,1]$ onto hyperinterval $D$ from (\ref{D}), a multidimensional problem (\ref{problemN}) can be reduced to a one-dimensional one:
 	\[
@@
 	\]
-It is known \cite{Strongin2000, Sergeyev2013} that the reduced one-dimensional function $\phi(y(x))$ will no longer be Lipschitzian, but will satisfy the H{\"o}lder condition
+	It is known \cite{Strongin2000, Sergeyev2013} that the reduced one-dimensional function $\phi(y(x))$ will no longer be Lipschitzian, but will satisfy the H{\"o}lder condition:
 	\[
@@
 
-Specific methods of numerical construction of this kind of mappings (called evolvents) are discussed in \cite{Strongin2000,Sergeyev2013}.
-Here we note that evolvent will approximate the theoretical Peano-Hilbert curve with an accuracy of $2^{-m}$, where $m$ is the parameter that sets the approximation accuracy.
-Examples of evolvent for different dimensions are shown in Fig.~\ref{fig:Peano}.
+	Specific methods for the numerical construction of this kind of mapping (called evolvents) are discussed in \cite{Strongin2000,Sergeyev2013}.
+	Here, we note that evolvent will approximate the theoretical Peano--Hilbert curve with an accuracy of $2^{-m}$, where $m$ is the parameter that sets the approximation accuracy.
+Examples of evolvent for different dimensions are shown in Figure~\ref{fig:Peano}.
 
-\begin{figure}
-\center
+	\begin{figure}[H]
+		\hspace{-9pt} 
 		\begin{minipage}{0.45\linewidth}
-\center{\includegraphics[width=0.9\linewidth]{fig1a.JPG} \\ (a)}
+			\center{\includegraphics[width=0.9\linewidth]{fig1a.JPG} \\ (\textbf{a})}
 		\end{minipage}
 		\begin{minipage}{0.45\linewidth}
-\center{\includegraphics[width=0.9\linewidth]{fig1b.JPG} \\ (b)}
+			\center{\includegraphics[width=0.9\linewidth]{fig1b.JPG} \\ (\textbf{b})}
 		\end{minipage}
-\caption{Evolvents with $m=4$ and (a) $N=2$, (b) $N=3$.}\label{fig:Peano}
+		\caption{Evolvents with $m=4$ and (\textbf{a}) $N=2$, (\textbf{b}) $N=3$.}\label{fig:Peano}
 	\end{figure}
@@
 	Thus, under this approach, the search trial at some point $x^k\in[0,1]$ will consist of constructing the image $y^k=y(x^k)$ first, and only then of calculating the function value $ z^k = \phi(y^k)$.
-We will call the three values $\{x^k,y ^k,z^k\}$ the trial result.
+	We will call the three values, $\{x^k,y ^k,z^k\}$, the trial result.
 
-\subsection{Asynchronous parallel algorithm used to determine the global minimum of a function}
+	\subsection{Asynchronous Parallel Algorithm Used to Determine the Global Minimum of a Function}
 
@@
 The master process performs a global search algorithm, in which the search information is accumulated.
-Then the Lipschitz constants for the objective function are estimated on its basis, and the points of new trials are calculated and distributed over the work processes.
+Then the Lipschitz constants for the objective function are estimated on this basis, and the points of new trials are calculated and distributed over the work processes.
 Worker-processes receive points from the master, carry out new trials in them, and send the results of the trials to the master.
@@
 	Let us assume that at each iteration, the master calculates one new trial  point and transfers it for processing to the worker.
-At the same time, a trial performed by a worker is a much more labor-intensive operation than selection of a new point by a master, which eliminates the worker downtime.
-In this case, the total number of trials performed by each worker  will depend on the labor intensity of the specific trial and cannot be estimated in advance.
+At the same time, a trial performed by a worker is a much more labor-intensive operation than the selection of a new point by a master, which eliminates the worker downtime.
+	In this case, the total number of trials performed by each worker  will depend on the labor intensity of the specific trial and cannot be estimated in~advance.
 
 	When describing a parallel algorithm, let us assume that there is one process master and $p$ workers.
-The upper index will indicate the serial number of the trial performed, and the lower index will be used to number the trials in ascending order of the coordinate.
+The {superscript} will indicate the serial number of the trial performed, and the {subscript} will be used to number the trials in ascending order of the coordinate.
 
@@
 The master process (let us assume that this is the process No 0) initiates  $p$ parallel trials at $p$ different points of the search region, two of which are boundary points, and the rest are arbitrary internal points (for example, uniformly distributed).
-Thus, the initial trials are carried out at the points $\{y(x^1), y(x^2), ...,y(x^p)\}$, where $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.
+Thus, the initial trials are carried out at the points $\{y(x^1), y(x^2), \dots,y(x^p)\}$, where $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,\dots, p-1$.
 
-Let us suppose that the master-process has obtained the results of $k$ trials, and at the current time the worker-processes are conducting trials at the points
-$\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.
+	Let us suppose that the master process has obtained the results of $k$ trials, and at the current time the worker processes are conducting trials at the points
+	$\{y(x^{k+1}), y(x^{k+2}), \dots,$ $y(x^{k+p})\}$.
 
-Let one of the processes (without limiting the commonality, we can assume that this is the process No 1), at some point in time complete the trial at the point $y(x^{k+1})$.
-The remaining processes are at the stage of conducting their trials, i.e. at these points multiple trials have already initiated, but have not yet been finalized.
+	Let one of the processes (without limiting the commonality, we can assume that this is the process No 1), at some point in time, complete the trial at the point $y(x^{k+1})$.
+	The remaining processes are at the stage of conducting their trials, i.e., at these points multiple trials have already been initiated, but have not yet been finalized.
 
-Then the No 1 process forwards to the master process the trial results, i.e. the three values $\{x^{k+1},y^{k+1},z^{k+1}\}$.
+	Then the No 1 process forwards to the master process the trial results, i.e., the three values $\{x^{k+1},y^{k+1},z^{k+1}\}$.
 	The master saves the results in its information base and selects a new trial point $x^{k+p+1}$ for the process No 1 in accordance with the rules described below.
 
-Step 1. Renumber he point of the set 
+	Step 1. Renumber {t}he points of the set 
 	\[
-X_k = \left\{x^1, x^2,...,x^{k+p} \right\},
+	X_k = \left\{x^1, x^2,\dots,x^{k+p} \right\},
 	\]
-which contains all the points at which the trials either have been carried out or are being carried out by the lower index in ascending order, i.e.
+	which contains all the points at which the trials have either been carried out or are being carried out by the lower index in ascending order, i.e.,
 	\[
-0=x_1<x_2<...<x_{x+p}=1.
+	0=x_1<x_2<\dots<x_{x+p}=1.
 	\]
+
 	Step 2. Calculate values
@@
 
-Step 3. For each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, calculate the value 
+	Step 3. For each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, calculate the~value 
 	\[
@@
 	\]
-called the characteristic of this interval. Here $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.
+	called the characteristic of this interval. Here, $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.
 
-Step 4. Select the interval $[x_{t-1},x_t]$ with the highest characteristic value, i.e.
+	Step 4. Select the interval $[x_{t-1},x_t]$ with the highest characteristic value, i.e.,
 	\[
@@
 
-Step 5. Calculate the point $x^{k+p+1} \in (x_{t-1},x_t)$ according to the formula
+	Step 5. Calculate the point $x^{k+p+1} \in (x_{t-1},x_t)$ according to the formula:
 	\[
@@
 
-The master stops the algorithm if either the condition $\Delta_{t}<\varepsilon$ (where $t$ is number of the interval with the maximum characteristic from Step 4) or the condition $k>K_{max}$ (where $k$ is the number of trials performed) is met.
+	The master stops the algorithm if either the condition $\Delta_{t}<\varepsilon$ ({where $t$ is the number} of the interval with the maximum characteristic from Step 4) or the condition $k>K_{max}$ (where $k$ is the number of trials performed) is met.
 	The real number $\varepsilon>0$ and the integer $K_{max} \geq 1$ are the algorithm parameters which correspond to the accuracy of the solution search and the maximum number of trials.
@@
 	The sequential global search algorithm, on the basis of which the proposed asynchronous algorithm was implemented, allows the following interpretation \cite{Molinaro2001}.
-Using the points of previously performed trials, it is possible to construct a minorant of the objective function, for which the value $-R(i)$, where $R(i)$ is the value of the characteristic of the interval $(x_{i-1}, x_i)$, will be the value of the minorant at the point of its minimum at the interval $(x_{i-1},x_i), 1\leq i\leq k+p$.
+Using the points of previously performed trials, it is possible to construct a minorant of the objective function, for which the value $-R(i)$, where $R(i)$ is the value of the {characteristics} of the interval $(x_{i-1}, x_i)$, will be the value of the minorant at the point of its minimum at the interval $(x_{i-1},x_i), 1\leq i\leq k+p$.
 In accordance with Step 4 of the algorithm, a new trial is carried out in the interval with the lowest value of the minorant.
-And in accordance with Step 5, the point of the new trial coincides with the minorant minimum point.
+In accordance with Step 5, the point of the new trial coincides with the minorant minimum point.
 
 	The theory of convergence and various modifications of serial and parallel algorithms are detailed in the monographs \cite{Strongin2000,Sergeyev2017}. 
-It should be noted that the asynchronous parallelization scheme described here, in contrast to the synchronous schemes used earlier in solving a number of application problems (see \cite{Kalyulin2017, Modorskii2016}), provides a full load of all the processes involved in solving problems with different labor intensity of trials at different points in the search domain.
+	It should be noted that the asynchronous parallelization scheme described here, in contrast to the synchronous schemes used earlier in solving a number of applied problems (see, e.g., \cite{Kalyulin2017}), provides a full load of all the processes involved in solving problems with different labor intensities of trials at different points in the search domain.
 
-\subsection{Features of the software implementation of the parallel algorithm}
-
-\begin{figure}
-\center{\includegraphics[width=0.9\linewidth]{fig2.JPG}}
-\caption{Implemented scheme of parallel computations (1 -- link via MPI, 2 -- link via streams).}\label{fig:Impl}
-\end{figure}
+	\subsection{Features of the Software Implementation of the Parallel Algorithm}
 
 	In our implementation of the global search algorithm, MPI is used to organize parallel calculations.
-A process with the rank 0 acts as a master, processes with a rank other than 0 act as calculators (workers), see the description of the algorithm in the previous section.
+A process with the rank 0 acts as a master, processes with a rank other than 0 act as calculators (workers) (see the description of the algorithm in the previous section).
 
 	The principle of process interaction (Figure \ref{fig:Impl}, link 1) is as follows.
-The first step in the master-process is to select $p$ points for trials.
+The first step in the master process is to select $p$ points for trials.
 The selected trial points are sent to the calculators and the master enters the stand-by state for receiving messages.
@@
 The new point is sent as a message to the calculator.
-If the stopping condition is met in the algorithm, instead of the trial point the master sends a shutdown message to the calculator.
+If the stopping condition is met in the algorithm, instead of the trial point, the master sends a shutdown message to the calculator.
+	\begin{figure}[H]
+		{\includegraphics[width=0.9\linewidth]{fig2.JPG}}
+		\caption{Implemented scheme of parallel computations (1---link via MPI, 2---link via streams).}\label{fig:Impl}
+	\end{figure}
 
-Each process that performs the role of a calculator creates an instance of the task when it starts, and then goes into a stand-by mode for messages from the master-process.
+	Each process that performs the role of a calculator creates an instance of the task when it starts, and then goes into a stand-by mode for messages from the master process.
 When a message containing a trial point is received, a computational experiment is performed.
@@
 
-When implementing the asynchronous scheme, we assumed that the start time of calculating the objective function values is different for each particular process.
+	When implementing the asynchronous scheme, we assumed that the start time for calculating the objective function values is different for each particular process.
 Our implementation takes into account the fact that the trial point as well as the calculation results represent a small amount of data that can be sent relatively quickly over the network.
@@
 First, the source code for the computational experiment was compiled into an executable application.
-Second, the approach in which MATLAB receives a trial point through input stream, and returns the results of computational experiment through output stream was implemented.
-This approach allowed each calculator to run a process with compiled MATLAB code once, intercept input and output streams, and then perform calculations without waiting for new processes to start.
+Second, the approach in which MATLAB received a trial point through input stream and returned the results of the computational experiment through output stream was implemented.
+This approach allowed each calculator to run a process with the compiled MATLAB code once, intercept input and output streams, and then perform calculations without waiting for new processes to start.
 
@@
 
-The asynchronous parallel global search algorithm was implemented on C++; the software was built on Intel MPI and Intel C++ compiler, version 2017.0.1.
+	The asynchronous parallel global search algorithm was implemented on C++; the software was built on Intel MPI and \hl{Intel C++ compiler, version 2017.0.1.} %MDPI: Please state the software creator, and the location (city, State Abbr., country) from where the software was sourced for ALL software in this paper.
  The objective function of the problem to be solved was implemented in MATLAB 2019b.
@@
 
-To solve the optimization problem, a combination of the parallel global search algorithm and a local optimization method was used.
-In the first stage of the search, the parallel asynchronous algorithm described in Section \ref{sec_GSA} was used.
-101 processes were involved in the launch, of which one process was a master and the remaining 100 were workers (calculators).
+	To solve the optimization problem, a combination of the parallel global search algorithm and a local optimization method was used. In the first stage of the search, 
+	{the parallel asynchronous algorithm described in Section \ref{sec_GSA} was used.
+The core parallel global search algorithm is available on GitHub, \url{https://github.com/UNN-ITMM-Software/gsa_nlp_solver} \hl{(accessed on).}} %MDPI: Please add the access date (Format: Date Month Year).
+e.g., (accessed on 1 January 2020).
+	One hundred and one processes were involved in the launch, of which one process was a master and the remaining 100 were workers (calculators).
 The values of the method parameters were set as follows: the maximum number of trials $K_{max}=3000$, the search accuracy $\varepsilon=10^{-3}$, the reliability parameter $r=3$, the evolvent accuracy parameter $m=12$.
-At the second stage, the solution was refined using the pattern search method with an accuracy of $\varepsilon=10^{-4}$ and a limit on the number of trials $K_{max}=500$.
-The total time for solving the problem was 8109 seconds, an optimum point was found with the objective function value $\phi^* = 0.002115$.
-Thus, using the developed algorithm, the inverse problem of chemical kinetics was solved and the kinetic parameters of the reaction were found, which are given in the table \ref{tab_kinetics}.
+At the second stage, the solution was refined
+	{using the Hooke--Jeeves pattern search method (see, e.g., \cite{HookJeeves})}
+	with an accuracy of $\varepsilon=10^{-4}$ and a limit on the number of trials $K_{max}=500$.
+	The total time for solving the problem was 8109 s and an optimum point was found with the objective function value $\phi^* = 0.002115$.
+Thus, using the developed algorithm, the inverse problem of chemical kinetics was solved and the kinetic parameters of the reaction were found, which are given in Table \ref{tab_kinetics}.
 
@@
 
-After the development of the kinetic reaction model, it is necessary to check the adequacy of the model, this procedure can be carried out using a direct kinetic problem.
-Substituting the found kinetic values $k_j^0$ and $E_j$ into the system \eqref{eq_concentrations} -- \eqref{eq_flow}, the direct problem of chemical kinetics was solved and the concentrations of reaction substances $x_i$ and temperature $T$ were found (Fig.
-\ref{fig:res1} -- \ref{fig:res2}).
+	After the development of the kinetic reaction model, it is necessary to check the adequacy of the model; this procedure can be carried out using a direct kinetic problem.
+Substituting the found kinetic values $k_j^0$ and $E_j$ into the system \eqref{eq_concentrations}--\eqref{eq_flow}, the direct problem of chemical kinetics was solved and the concentrations of reaction substances $x_i$ and temperature $T$ were found (Figures \ref{fig:res1} and \ref{fig:res2}).
 
-Figures \ref{fig:res1} -- \ref{fig:res2} show graphs of changes in the concentrations of individual components of the reaction and temperature change in the reactor unit.
+	Figures \ref{fig:res1} and \ref{fig:res2} show graphs of changes in the concentrations of individual components of the reaction and temperature changes in the reactor unit.
 It should be noted that in the figures the vertical lines separating the graphs characterize three consecutive reactors (I, II, III).
@@
 
-\begin{figure}
-\center
-\begin{minipage}{0.45\linewidth}
-\center{\includegraphics[width=0.9\linewidth]{1.eps} \\ (a)}
+	\begin{figure}[H]
+		\hspace{-12pt}
+		\begin{minipage}{0.5\linewidth}
+			\center{\includegraphics[width=0.9\linewidth]{1.eps} \\ (\textbf{a})}
 		\end{minipage}
-\begin{minipage}{0.45\linewidth}
-\center{\includegraphics[width=0.9\linewidth]{2.eps} \\ (b)}
+		\begin{minipage}{0.5\linewidth}
+			\center{\includegraphics[width=0.9\linewidth]{2.eps} \\ (\textbf{b})}
 		\end{minipage}
-\caption{Changes in concentrations of (a) $n$-pentane ($n$-C$_5$) and $n$-hexane ($n$-C$_6$), (b) isopentane (i-C5), 2-methylpentane (2-MP), 3-methylpentane (3-MP), 2,2-dimethylbutane(2,2-DMB), 2,3-dimethylbutane (2,3-DMB).}\label{fig:res1}
+		\caption{\hl{Changes} %MDPI: Please use commas to separate thousands for numbers with five or more digits (not for four digits) in the picture. e.g., "10000" should be "10,000"
+ in concentrations of (\textbf{a}) $n$-pentane ($n$-C$_5$) and $n$-hexane ($n$-C$_6$), (\textbf{b}) isopentane (i-C5), 2-methylpentane (2-MP), 3-methylpentane (3-MP), 2,2-dimethylbutane(2,2-DMB), and 2,3-dimethylbutane (2,3-DMB).}\label{fig:res1}
+	\end{figure}  
+	\vspace{-12pt} 
+	\begin{figure}[H]
+		\hspace{-15pt} 	
+		\begin{minipage}{0.5\linewidth}
+		\center{\includegraphics[width=0.9\linewidth]{3.eps} \\ (\textbf{a})}
+		\end{minipage}
+		\begin{minipage}{0.5\linewidth}
+			\center{\includegraphics[width=0.9\linewidth]{4.eps} \\ (\textbf{b})}
+		\end{minipage}
+		\caption{\hl{Changes} 
+		%MDPI: (1) Please change the hyphen (-) into minus sign ($-$, "U+2212").
+e.g., "-1" should be "$-$1".
+(2) Please use commas to separate thousands for numbers with five or more digits (not for four digits) in the picture.
+e.g., "10000" should be "10,000".
+(3)  Please add the explanation for BLUE BOX in the figure.
+ in (\textbf{a}) concentrations of benzene, cyclohexane (CH*), methyl cyclopentane (MCP), and cyclopentane (CP), (\textbf{b}) temperature.}\label{fig:res2}
 	\end{figure}
 
-\begin{figure}
-\center
-\begin{minipage}{0.45\linewidth}
-\center{\includegraphics[width=0.9\linewidth]{3.eps} \\ (a)}
-\end{minipage}
-\begin{minipage}{0.45\linewidth}
-\center{\includegraphics[width=0.9\linewidth]{4.eps} \\ (b)}
-\end{minipage}
-\caption{Changes in (a) concentrations of benzene, cyclohexane (CH*), methyl cyclopentane (MCP), cyclopentane (CP), (b) temperature.}\label{fig:res2}
-\end{figure}
+	Figure \ref{fig:res1}a shows that the concentrations of n-pentane and n-hexane are continuously decreasing.
+This is mainly due to the transformation of the alkanes with a normal (linear) structure into the corresponding isomers, which is reflected in Figure \ref{fig:res1}b.
+With the transformation of n-pentane, everything is natural, since only one isomer is known.
+In the case of n-hexane, everything is somewhat different: first, the conversion of n-hexane into isomers 2-methylpentane and 3-methylpentane is carried out and the latter are further converted into 2,2-dimethylbutane and 2,3-dimethylbutane.
 
-Figure \ref{fig:res1} (a) shows that the concentrations of n-pentane and n-hexane are continuously decreasing.
-This is mainly due to the transformation of alkanes of normal (linear) structure into the corresponding isomers, which is reflected in Figure \ref{fig:res1} (b).
-With the transformation of n-pentane, everything is natural, since only one isomer is known.
-And in the case of n-hexane, everything is somewhat different: first, the conversion of n-hexane is carried out into isomers 2-methylpentane and 3-methylpentane, and the latter are further converted into 2,2-dimethylbutane and 2,3-dimethylbutane.
+	In Figure \ref{fig:res2}a, it can be seen that the concentration of cyclopentane in reactor I gradually decreases to zero.
+The decrease in cyclopentane concentration can be explained by the hydrogenation reaction to $n$-pentane.
+According to Figure \ref{fig:res1}a, this is probably the reason for the increase in the concentration of $n$-pentane at the beginning of reactor I.
+	Additionally, this figure shows a sharp decrease in the concentration of benzene, presumably due to hydrogenation reactions.
 
-According to Figure \ref{fig:res2} (a), it can be seen that the concentration of cyclopentane in reactor I gradually decreases to zero.
-The decrease in cyclopentane concentration can be explained by the hydrogenation reaction to $n$-pentane.
-According to Figure \ref{fig:res1} (a), this is probably the reason for the increase in the concentration of $n$-pentane at the beginning of reactor I.
-Also, this figure shows a sharp decrease in the concentration of benzene, presumably due to hydrogenation reactions.
-
-Figure \ref{fig:res2} (b) shows a graph of temperature changes in reactors I-III.
-According to the technological scheme, refrigeration equipment for cooling the reaction mixture is installed after reactor II.
+	Figure \ref{fig:res2}b shows a graph of temperature changes in reactors I--III.
+According to the technological scheme, refrigeration equipment for cooling the reaction mixture is installed after reactor II; {therefore, a sharp decrease in temperature is observed when the catalyst is 18,000 kg}.
 A gradual increase in temperature in reactors III here indicates the presence of isomerization reactions that occur with a slight exothermic effect.
 
-It is worth noting that the composition of the products at the outlet of the last reactor and the temperature differences in all reactors are close to industrial data.
+	It is worth noting that the composition of the products at the outlet of the last reactor and the temperature differences in all reactors are close to the industrial data.
 Thus, these facts may indicate the adequacy of the model we have developed.
+
+	\section{Conclusions}\label{sec_conclusion}
+
+	\hl{To} %MDPI: We have removed the color, please confirm.
+ summarize, the problem of developing a kinetic model for the process of catalytic isomerization of the pentane-hexane fraction using a parallel optimization method was considered.
+Due to the switch of the oil refining industry to the production of gasoline with  the Euro-5 standard, this process is becoming important.
+On the basis of known industrial data on the concentrations of reaction components and the temperature at the outlet of the third reactor, the inverse problem of chemical kinetics was solved, and the activation energies and pre-exponential factors of each reaction stage were calculated.
+To solve the optimization problem, the authors have developed a parallel global search algorithm and a program based on Lipschitz optimization.
+The kinetic parameters found made it possible to develop a mathematical model of the process, which is in good agreement with industrial data.
+
+	{The future plan is to study the dynamics of the gas--liquid flow in the reactor unit for the process of catalytic isomerization of the pentane-hexane fraction, taking into account diffusion and heat exchange processes through the catalyst layer.}
 
@@
 and A.F.; writing---review and editing, R.F.
-and L.E.; visualization, R.F.; supervision, I.G..
+and L.E.; visualization, R.F.; supervision, I.G.
 All authors have read and agreed to the published version of the manuscript.}
 
-\funding{This research was funded by RFBR, project No.\,19-37-60014 (L.E.; mathematical model), and by the Russian Science Foundation, project No.\,21-11-00204 (K.B, I.L., E.K.; parallel algorithm)}
+	\funding{\hl{This} %MDPI: Information regarding the funder and the funding number should be provided. Please check the accuracy of funding data and any other information carefully.
+ research was funded by RFBR, project No.\,19-37-60014 (L.E.; mathematical model), and by the Russian Science Foundation, project No.\,21-11-00204 (K.B, I.L., E.K.; parallel global optimization algorithm)}
 
@@
 	%% Optional
-\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+	\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A").
+If the appendix sections contain a heading then change the argument to "yes".
 	\appendixstart
@@
 	\begin{table}[H] 
-\caption{Kinetic parameters of the catalytic isomerisation reaction of the pentane-hexane fraction, according to the scheme of chemical transformations (Table \ref{tab_chemistry}).\label{tab_kinetics}}
+		\caption{\hl{Kinetic} %MDPI: Please confirm if the italics should be retained.
+ parameters of the catalytic isomerisation reaction of the pentane-hexane fraction, according to the scheme of chemical transformations (Table \ref{tab_chemistry}).\label{tab_kinetics}}
 		\newcolumntype{C}{>{\centering\arraybackslash}X}
@@
 			\toprule
-\textbf{No}	& \textbf{$E$, kcal/mol} & \textbf{$\log k_0$} & \textbf{No}	& \textbf{$E$, kcal/mol} & \textbf{$\log k_0$} \\
+			\textbf{No}	& \textbf{\boldmath{$E$}, kcal/mol} & \boldmath{$\log k_0$} & \textbf{No}	& \textbf{\boldmath{$E$}, kcal/mol} & \boldmath{$\log k_0$} \\
 			\midrule
@@
 			6  & 162.603 & 17.997 & 30 & 97.215 & 9.612 \\			
+			\bottomrule
+		\end{tabularx}
+	\end{table}
+
+\begin{table}[H]\ContinuedFloat
+\caption{{\em Cont.}}
+	\newcolumntype{C}{>{\centering\arraybackslash}X}
+		\begin{tabularx}{\textwidth}{CCCCCC}
+			\toprule
+			\textbf{No}	& \textbf{\boldmath{$E$}, kcal/mol} & \boldmath{$\log k_0$} & \textbf{No}	& \textbf{\boldmath{$E$}, kcal/mol} & \boldmath{$\log k_0$} \\
+			\midrule
 			7 & 161.429 & 17.114 & 31 & 188.306 & 7.194  \\
@@
 		%=====================================
-\bibliography{bibliography.bib}
+		\begin{thebibliography}{999}
+
+\bibitem[Weyda and Köhler(2003)]{Hartmut}
+Weyda, H.; Köhler, E.
+\newblock Modern refining concepts—An update on naphtha-isomerization to
+  modern gasoline manufacture.
+\newblock {\em Catal. Today} {\bf 2003}, {\em 81},~51--55.
+
+\bibitem[Fahim \em{et~al.}(2010)Fahim, Alsahhaf, and Elkilani]{Fahim}
+Fahim, M.; Alsahhaf, T.; Elkilani, A.
+\newblock {\em Fundamentals of Petroleum Refining}; Elsevier:  \hl{Amsterdam, The Netherlands,} %MDPI: We added the location of publisher. Please confirm
+ 2010.
+
+\bibitem[Meyers(1997)]{Meyers}
+Meyers, R.A.
+\newblock {\em Handbook of Petroleum Refining Processes}; McGraw-Hill: \hl{New York, NY, USA},  1997.
+
+\bibitem[Dao \em{et~al.}(2022)Dao, Tran, Tran, and Lam]{DAO2022172}
+Dao, T.P.; Tran, N.Q.; Tran, T.T.; Lam, V.T.
+\newblock Assessing the kinetic model on extraction of essential oil and
+  chemical composition from lemon peels (\emph{Citrus aurantifolia}) by
+  hydro-distillation process.
+\newblock {\em Mater. Today Proc.} {\bf 2022}, {\em 51},~172--177.
+
+\bibitem[Sajid \em{et~al.}(2020)Sajid, Bai, Liu, and Zhao]{SAJID20207430}
+Sajid, M.; Bai, Y.; Liu, D.; Zhao, X.
+\newblock Organic acid catalyzed production of platform chemical
+  5-hydroxymethylfurfural from fructose: Process comparison and evaluation
+  based on kinetic modeling.
+\newblock {\em Arab. J. Chem.} {\bf 2020}, {\em 13},~7430--7444.
+
+\bibitem[Li and Cheng(2021)]{LI2021175}
+Li, Y.; Cheng, H.
+\newblock Chemical kinetic modeling of organic pollutant degradation in Fenton
+  and solar photo-Fenton processes.
+\newblock {\em J. Taiwan Inst. Chem. Eng.} {\bf
+  2021}, {\em 123},~175--184.
+
+\bibitem[Hou \em{et~al.}(2017)Hou, Gao, Li, Zhang, Wang, Song, and
+  Yao]{HOU201759}
+Hou, L.; Gao, K.; Li, P.; Zhang, X.; Wang, Z.; Song, P.; Yao, W.
+\newblock A kinetic model for calculating total membrane fouling resistance in
+  chemical cleaning process.
+\newblock {\em Chem. Eng. Res. Des.} {\bf 2017}, {\em
+  128},~59--72.
+
+\bibitem[Uskov \em{et~al.}(2019)Uskov, Potemkin, Kamboj, Snytnikov,
+  Pakharukova, Enikeeva, Gubaydullin, and Hussainova]{uskovfibrous}
+Uskov, S.; Potemkin, D.; Kamboj, N.; Snytnikov, P.; Pakharukova, V.; Enikeeva,
+  L.; Gubaydullin, I.; Hussainova, I.
+\newblock Fibrous alumina-based Ni-MOx (M=Mg, Cr, Ce) catalysts for propane
+  pre-reforming.
+\newblock {\em Mater. Lett.} {\bf 2019}, {\em 257},~126741.
+\newblock {{https://doi.org/10.1016/j.matlet.2019.126741}}.
+
+\bibitem[Uskov \em{et~al.}(2020)Uskov, Potemkin, Enikeeva, Snytnikov,
+  Gubaydullin, and Sobyanin]{en13133393}
+Uskov, S.I.; Potemkin, D.I.; Enikeeva, L.V.; Snytnikov, P.V.; Gubaydullin,
+  I.M.; Sobyanin, V.A.
+\newblock Propane Pre-Reforming into Methane-Rich Gas over Ni Catalyst:
+  Experiment and Kinetics Elucidation via Genetic Algorithm.
+\newblock {\em Energies} {\bf 2020}, {\em 13}, 3393.
+\newblock {{https://doi.org/10.3390/en13133393}}.
+
+\bibitem[Enikeeva \em{et~al.}(2021)Enikeeva, Faskhutdinov, Koledina,
+  Faskhutdinova, and Gubaydullin]{enik2021}
+Enikeeva, L.; Faskhutdinov, A.; Koledina, K.; Faskhutdinova, R.; Gubaydullin,
+  I.
+\newblock Modeling and optimization of the catalytic isomerization of the
+  pentane-hexane fraction with maximization of individual high-octane
+  components yield.
+\newblock {\em React. Kinet. Mech. Catal.} {\bf 2021}, {\em
+  133},~879--895.
+\newblock {{https://doi.org/10.1007/s11144-021-02020-w}}.
+
+\bibitem[Enikeeva \em{et~al.}(2020)Enikeeva, Marchenko, Smirnov, and
+  Gubaydullin]{Enikeeva2020}
+Enikeeva, L.; Marchenko, M.; Smirnov, D.; Gubaydullin, I.
+\newblock Parallel Gravitational Search Algorithm in Solving the Inverse
+  Problem of Chemical Kinetics.
+\newblock In \emph{Proceedings of the Supercomputing}; Voevodin, V., Sobolev, S.,
+  Eds.; Springer International Publishing: Cham, Switzerland, 2020; pp. 98--109.
+\newblock {{https://doi.org/10.1007/978-3-030-64616-5\_9}}.
+
+\bibitem[Paulavi{\v c}ius and {\v Z}ilinskas(2014)]{PaulaviciusZilinskas2014}
+Paulavi{\v c}ius, R.; {\v Z}ilinskas, J.
+\newblock {\em Simplicial Global Optimization}; Springer: New York, NY, USA,  2014.
+\newblock {{https://doi.org/10.1007/978-1-4614-9093-7}}.
+
+\bibitem[Sergeyev and Kvasov(2017)]{Sergeyev2017}
+Sergeyev, Y.D.; Kvasov, D.E.
+\newblock {\em Deterministic Global Optimization: An Introduction to the
+  Diagonal Approach}; Springer: New York, NY, USA,  2017.
+\newblock {{https://doi.org/10.1007/978-1-4939-7199-2}}.
+
+\bibitem[Battiti \em{et~al.}(2009)Battiti, Brunato, and Mascia]{Battiti2009}
+Battiti, R.; Brunato, M.; Mascia, F.
+\newblock {\em Reactive Search and Intelligent Optimization}; Springer: New
+  York, NY, USA,  2009.
+\newblock {{https://doi.org/10.1007/978-0-387-09624-7}}.
+
+\bibitem[Eiben and Smith(2015)]{Eiben2015}
+Eiben, A.; Smith, J.
+\newblock {\em Introduction to Evolutionary Computing}; Springer:  \hl{Berlin, Germany,} %MDPI: We added the location of publisher. Please confirm
+  2015.
+\newblock {{https://doi.org/10.1007/978-3-662-44874-8}}.
+
+\bibitem[Kvasov and Mukhametzhanov(2018)]{Kvasov2018}
+Kvasov, D.; Mukhametzhanov, M.
+\newblock Metaheuristic vs. deterministic global optimization algorithms: The
+  univariate case.
+\newblock {\em Appl. Math. Comput.} {\bf 2018}, {\em 318},~245--259.
+\newblock {{https://doi.org/10.1016/j.amc.2017.05.014}}.
+
+\bibitem[Sergeyev \em{et~al.}(2018)Sergeyev, Kvasov, and
+  Mukhametzhanov]{Sergeyev2018}
+Sergeyev, Y.; Kvasov, D.; Mukhametzhanov, M.
+\newblock On the efficiency of nature-inspired metaheuristics in expensive
+  global optimization with limited budget.
+\newblock {\em Sci. Rep.} {\bf 2018}, {\em 8},~435.
+
+\bibitem[Sovrasov(2019)]{Sovrasov2019}
+Sovrasov, V.
+\newblock Comparison of several stochastic and deterministic derivative-free
+  global optimization algorithms.
+\newblock {\em Lect. Notes Comput. Sci.} {\bf 2019}, {\em
+  11548},~70--81.
+
+\bibitem[Barkalov and Lebedev(2016)]{Barkalov2016}
+Barkalov, K.; Lebedev, I.
+\newblock Solving multidimensional global optimization problems using graphics
+  accelerators.
+\newblock {\em Commun. Comput. Inf. Sci.} {\bf 2016},
+  {\em 687},~224--235.
+
+\bibitem[Strongin \em{et~al.}(2018)Strongin, Gergel, Barkalov, and
+  Sysoyev]{Strongin2018}
+Strongin, R.; Gergel, V.; Barkalov, K.; Sysoyev, A.
+\newblock Generalized Parallel Computational Schemes for Time-Consuming Global
+  Optimization.
+\newblock {\em Lobachevskii J. Math.} {\bf 2018}, {\em
+  39},~576--586.
+
+\bibitem[Sysoyev \em{et~al.}(2017)Sysoyev, Barkalov, Sovrasov, Lebedev, and
+  Gergel]{Sysoyev2017}
+Sysoyev, A.; Barkalov, K.; Sovrasov, V.; Lebedev, I.; Gergel, V.
+\newblock Globalizer---A parallel software system for solving global
+  optimization problems.
+\newblock {\em Lect. Notes Comput. Sci.} {\bf 2017}, {\em
+  10421},~492--499.
+
+\bibitem[Logan(1996)]{logan}
+Logan, S.
+\newblock {\em Fundamentals of Chemical Kinetics}; Prentice Hall: \hl{Hoboken, NJ, USA},  1996.
+
+\bibitem[Kvasov and Sergeyev(2013)]{Kvasov2013}
+Kvasov, D.; Sergeyev, Y.
+\newblock Lipschitz global optimization methods in control problems.
+\newblock {\em Autom. Remote. Control.} {\bf 2013}, {\em 74},~1435--1448.
+\newblock {{https://doi.org/10.1134/S0005117913090014}}.
+
+\bibitem[Kalyulin \em{et~al.}(2017)Kalyulin, Shavrina, Modorskii, Barkalov, and
+  Gergel]{Kalyulin2017}
+Kalyulin, S.; Shavrina, E.; Modorskii, V.; Barkalov, K.; Gergel, V.
+\newblock Optimization of Drop Characteristics in a Carrier Cooled Gas Stream
+  Using \uppercase{ANSYS} and \uppercase{G}lobalizer Software Systems on the
+  \uppercase{PNRPU} High-Performance Cluster.
+\newblock {\em Commun. Comput. Inf. Sci.} {\bf 2017},
+  {\em 753},~331--345.
+\newblock {{https://doi.org/10.1007/978-3-319-67035-5\_24}}.
+
+\bibitem[{Strongin R.G., Sergeyev Y.D.}(2000)]{Strongin2000}
+{Strongin R.G.; Sergeyev Y.D.}
+\newblock {\em Global Optimization with Non-Convex Constraints. Sequential and
+  Parallel Algorithms}; Kluwer Academic Publishers: Dordrecht, The Netherlands, 2000.
+\newblock {{https://doi.org/10.1007/978-1-4615-4677-1}}.
+
+\bibitem[Jones \em{et~al.}(1993)Jones, Perttunen, and Stuckman]{Jones1993}
+Jones, D.; Perttunen, C.; Stuckman, B.
+\newblock Lipschitzian optimization without the \uppercase{L}ipschitz constant.
+\newblock {\em J. Optim. Theory Appl.} {\bf 1993}, {\em 79},~157--181.
+
+\bibitem[Grishagin \em{et~al.}(2018)Grishagin, Israfilov, and
+  Sergeyev]{Grishagin2018}
+Grishagin, V.; Israfilov, R.; Sergeyev, Y.
+\newblock Convergence conditions and numerical comparison of global
+  optimization methods based on dimensionality reduction schemes.
+\newblock {\em Appl. Math. Comput.} {\bf 2018}, {\em 318},~270--280.
+\newblock {{https://doi.org/10.1016/j.amc.2017.06.036}}.
+
+\bibitem[{Sergeyev, Y.D., Strongin, R.G., Lera, D.}(2013)]{Sergeyev2013}
+{Sergeyev, Y.D.; Strongin, R.G.; Lera, D}.
+\newblock {\em Introduction to Global Optimization Exploiting Space-Filling
+  Curves}; Springer Briefs in Optimization; Springer: New York, NY, USA, 2013.
+\newblock {{https://doi.org/10.1007/978-1-4614-8042-6}}.
+
+\bibitem[Molinaro \em{et~al.}(2001)Molinaro, Pizzuti, and
+  Sergeyev]{Molinaro2001}
+Molinaro, A.; Pizzuti, C.; Sergeyev, Y.
+\newblock Acceleration tools for diagonal information global optimization
+  algorithms.
+\newblock {\em Comput. Optim. Appl.} {\bf 2001}, {\em
+  18},~5--26.
+\newblock {{https://doi.org/10.1023/A:1008719926680}}.
+
+\bibitem[Hooke and Jeeves(1961)]{HookJeeves}
+Hooke, R.; Jeeves, T.
+\newblock ``\uppercase{D}irect Search'' Solution of Numerical and Statistical
+  Problems.
+\newblock {\em J. ACM} {\bf 1961}, {\em 8},~212--229.
+\newblock {{https://doi.org/10.1145/321062.321069}}.
+
+\end{thebibliography}
 

### 2022-10-11 2022/PAVT Special Issue/template.tex

@@
 \abstract{The article is devoted to the problem of developing a kinetic model of a complex chemical reaction using a parallel optimization method.
-The design of the kinetic model consists in finding the kinetic parameters of the reaction, which cannot be calculated analytically, and since the chemical reaction involves many stages, the optimization problem is multi-extreme.
+The design of the kinetic model consists in finding the kinetic parameters of the reaction, which cannot be calculated analytically, and since the chemical reaction involves many stages, the optimization problem is \textcolor[rgb]{1,0,0}{multiextremal}.
 As a chemical reaction, the process of catalytic isomerization of the pentane-hexane fraction is considered, which is now important due to the switch of the oil refining industry to the production of gasoline corresponding to the Euro-5 standard.
@@
 Furthermore, in the inverse problems of chemical kinetics, the objective function can be substantially multiextremal, i.e., having many local extremes along the global one.
+\textcolor[rgb]{1,0,0}{Gradient-based optimization methods will not give a good result for such problems.
+On the one hand, such methods converge only to a local solution.
+On the other hand, in the considered problem the gradient of the objective function is not known analytically and its numerical estimate will be time-consuming operation.}
 
@@
 It is assumed that the values of the objective function are determined as a result of solving some other task and require a costly computational experiment.
+\textcolor[rgb]{1,0,0}{On many benchmarks Lipschitz global optimization methods outperform metaheuristic algorithms in terms of the number of correctly solved problems \cite{Sergeyev2018,Sovrasov2019}.}
 
@@
 \dfrac{dT}{d\tau} = \dfrac{\sum_{i=1}^I\left( \dfrac{dx_i}{d\tau} \cdot \Delta H^i_{f,T} \right) }{\sum_i^J \left(x_i C^T_{p_i} \right) },
-\end{equation}\label{eq_flow}
-\begin{equation}
+\end{equation}
+\begin{equation}\label{eq_flow}
 \dfrac{dF}{d\tau} = \sum_{i=1}^I  \dfrac{x_i}{d\tau}
@@
 \begin{linenomath}
-\begin{equation}
+\begin{equation}\label{FF}
 FF = \dfrac{\sum_{i=1}^I \left| \dfrac{y_i^{exp} - y_i^{calc}}{y_i^{exp}} \right|}{I} \rightarrow \min,
@@
 \end{linenomath}
-
 where I is the number of observed parameters (I=19, of which 17 parameters are $x_i$, one parameter is $T$ and one parameter is $F$),  $y_{i}^{exp}$ are industrual data, $y_{i}^{calc}$ are numerical values of concentrations of reaction substances obtained by solving a direct kinetic problem.
@@
 As stated above, from a mathematical point of view, the inverse problem of chemical kinetics can be considered a Lipschitz global optimization problem.
+\textcolor[rgb]{1,0,0}{This fact has has the following justification.
+We solve the ODE system (\ref{eq_concentrations})--(\ref{eq_flow}) to compute the value of the objective function (\ref{FF}).
+The right-hand side of the ODE system is continuous functions with bounded derivatives.
+Thus, the system solution will also be continuous and bounded at the bounded period of time.
+Therefore, the objective function (\ref{FF}) will satisfy the Lipschitz condition, but the Lipschitz constant will be a priori unknown.}
+
 In general, the problem of the specified class can be formulated as follows:
@@
 The criterion $\phi(y)$ is assumed to be multiextremal and is given in the form of some computational algorithm. 
+\textcolor[rgb]{1,0,0}{This assumption is typical for many applied optimization problems (see, e.g., \cite{Kvasov2013,Kalyulin2017}).}
 The input of this algorithm is supplied with a parameter vector, and the output is the corresponding function value.
@@
 When describing a parallel algorithm, let us assume that there is one process master and $p$ workers.
-The upper index will indicate the serial number of the trial performed, and the lower index will be used to number the trials in ascending order of the coordinate.
+The \textcolor[rgb]{1,0,0}{superscript} will indicate the serial number of the trial performed, and the \textcolor[rgb]{1,0,0}{subscript} will be used to number the trials in ascending order of the coordinate.
 
@@
 
-Step 1. Renumber he point of the set 
+Step 1. Renumber \textcolor[rgb]{1,0,0}{t}he points of the set 
 \[
@@
 
-The master stops the algorithm if either the condition $\Delta_{t}<\varepsilon$ (where $t$ is number of the interval with the maximum characteristic from Step 4) or the condition $k>K_{max}$ (where $k$ is the number of trials performed) is met.
+The master stops the algorithm if either the condition $\Delta_{t}<\varepsilon$ (\textcolor[rgb]{1,0,0}{where $t$ is the number} of the interval with the maximum characteristic from Step 4) or the condition $k>K_{max}$ (where $k$ is the number of trials performed) is met.
 The real number $\varepsilon>0$ and the integer $K_{max} \geq 1$ are the algorithm parameters which correspond to the accuracy of the solution search and the maximum number of trials.
@@
 The sequential global search algorithm, on the basis of which the proposed asynchronous algorithm was implemented, allows the following interpretation \cite{Molinaro2001}.
-Using the points of previously performed trials, it is possible to construct a minorant of the objective function, for which the value $-R(i)$, where $R(i)$ is the value of the characteristic of the interval $(x_{i-1}, x_i)$, will be the value of the minorant at the point of its minimum at the interval $(x_{i-1},x_i), 1\leq i\leq k+p$.
+Using the points of previously performed trials, it is possible to construct a minorant of the objective function, for which the value $-R(i)$, where $R(i)$ is the value of the \textcolor[rgb]{1,0,0}{characteristics} of the interval $(x_{i-1}, x_i)$, will be the value of the minorant at the point of its minimum at the interval $(x_{i-1},x_i), 1\leq i\leq k+p$.
 In accordance with Step 4 of the algorithm, a new trial is carried out in the interval with the lowest value of the minorant.
@@
 
-To solve the optimization problem, a combination of the parallel global search algorithm and a local optimization method was used.
-In the first stage of the search, the parallel asynchronous algorithm described in Section \ref{sec_GSA} was used.
+To solve the optimization problem, a combination of the parallel global search algorithm and a local optimization method was used. In the first stage of the search, 
+\textcolor[rgb]{1,0,0}{the parallel asynchronous algorithm described in Section \ref{sec_GSA} was used.
+The core parallel global search algorithm is available on GitHub, \url{https://github.com/UNN-ITMM-Software/gsa_nlp_solver}.}
 101 processes were involved in the launch, of which one process was a master and the remaining 100 were workers (calculators).
 The values of the method parameters were set as follows: the maximum number of trials $K_{max}=3000$, the search accuracy $\varepsilon=10^{-3}$, the reliability parameter $r=3$, the evolvent accuracy parameter $m=12$.
-At the second stage, the solution was refined using the pattern search method with an accuracy of $\varepsilon=10^{-4}$ and a limit on the number of trials $K_{max}=500$.
+At the second stage, the solution was refined
+\textcolor[rgb]{1,0,0}{using the Hooke-Jeeves pattern search method (see, e.g., \cite{HookJeeves})}
+with an accuracy of $\varepsilon=10^{-4}$ and a limit on the number of trials $K_{max}=500$.
 The total time for solving the problem was 8109 seconds, an optimum point was found with the objective function value $\phi^* = 0.002115$.
@@
 
-\funding{This research was funded by RFBR, project No.\,19-37-60014 (L.E.; mathematical model), and by the Russian Science Foundation, project No.\,21-11-00204 (K.B, I.L., E.K.; parallel algorithm)}
+\funding{This research was funded by RFBR, project No.\,19-37-60014 (L.E.; mathematical model), and by the Russian Science Foundation, project No.\,21-11-00204 (K.B, I.L., E.K.; parallel global optimization algorithm)}
 

### 2022-10-11 2022/NumMethOneAPI/NMP_authorGuidelinesRu_2022.tex
check first section

@@
 %===========================  ШАПКА СТАТЬИ =================================
-\Article{Решение трудоемких задач многомерной глобальной оптимизации с использованием набора инструментов Intel oneAPI
-}
+\Article{О реализации параллельного алгоритма глобальной оптимизации с использованием набора инструментов Intel oneAPI
+}%Обновить перевод названия
     {Solving time-consuming multidimensional global optimization problems with Intel OneApi tools}
 
-\Abstract{В рамках данной статьи рассматривается решение серии сложно вычислимых задач глобальной оптимизации.
-В качестве целевой выбирается функция, удовлетворяющая условию Липшица с заранее не известной константой.
-Инструментом для реализации параллельного алгоритма выбран Intel oneAPI, который позволяет писать один код как для центрального процессора, так и для графического ускорителя.
-В решении многомерных задач применяется метод, построенный на идее редукции размерности при помощи кривой Пеано, которая непрерывно и однозначно отображает отрезок вещественной оси на гиперкуб.}
+\Abstract{В статье рассматривается параллельный алгоритм решения задач глобальной оптимизации и обсуждается его реализация с использованием набора инструментов Intel oneAPI.
+Предполагается, что целевая функция задачи задана как <<черный ящик>> и удовлетворяет условию Липшица.
+Изложенный в статье параллельный алгоритм использует схему редукции размерности на основе кривых Пеано, которые непрерывно и однозначно отображают отрезок вещественной оси на гиперкуб.
+В качестве средства для реализации параллельного алгоритма использован инструментарий Intel oneAPI, который позволяет писать один код как для центрального процессора, так и для графических ускорителей.
+Приведены результаты вычислительных экспериментов, полученные при решении серии сложных задач многоэкстремальной оптимизации.}
+%обновить перевод - сейчас он для старой аннотации
     {The paper considers the process of solving a series of time-consuming global optimization problems.
@@
 \VOL{??}{?}
-\YEAR{2021}
-\Received{16 июня 2021 г.}{June 16, 2021}
-\Accepted{13 октября 2021 г.}{October 13, 2021}
+\YEAR{2022}
+\Received{16 июня 2022 г.}{June 16, 2022}
+\Accepted{13 октября 2022 г.}{October 13, 2022}
 
@@
 Задачи поиска экстремумов многоэкстремальных функций часто возникают в различных областях науки, таких как химия, физика, машиностроение и т.д.
-В рамках данной работы решается задача минимизации многомерной функции $\varphi(y)$ в гиперинтервале \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\)  Будем предполагать, что информация о виде самой функции недоступна, целевая функция задается в формате «черного ящика».
-Дополнительно для целевой функции предполагаем выполнение условия Липшица, причем константа L априори неизвестна.
 
+В рамках данной работы рассматривается задача поиска глобального минимума $y^*$ функции $\varphi(y)$ в гиперинтервале $D$, т.е. 
 \begin{equation}
 \label{task}
-\varphi(y^*)=\min\{\varphi(y):y\in D\},
+\varphi(y^*)=\min\{\varphi(y):y\in D\}, \; D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}.
 \end{equation}
 
+Будем предполагать, что информация о структурных свойствах целевой функции $\varphi(y)$ неизвестна, а сама функция задается как <<черный ящик>>, т.е.
+с помощью некоторого реализованного программно алгоритма вычисления ее значений.
+Также будем предполагать, что $\varphi(y)$ удовлетворяет Липшица
 \begin{equation}
 \label{lip}
-|\varphi(y_1)-\phi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty.
+|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty,
 \end{equation}
+причем константа Липшица $L$ априори неизвестна.
 
-Сведение многомерной задачи ((\ref{task})) к одномерной проводится за счет использования редукции размерности при помощи кривой Пеано [\ref{rfa:rulit:lit1}, \ref{rfa:rulit:lit2}] y(x), которая непрерывно и однозначно отображает на N-мерный гиперкуб отрезок вещественной оси [0,1]:
+%комментарий про разные способы редукции размерности - ссылки на Сергеева, Жилинскаса
 
+Сведение многомерной задачи (\ref{task}) к одномерной проводится за счет использования редукции размерности при помощи кривой Пеано [\ref{rfa:rulit:lit1}, \ref{rfa:rulit:lit2}] $y(x)$, которая непрерывно и однозначно отображает отрезок вещественной оси $[0,1]$ на $N$-мерный гиперкуб:
 \begin{equation}
 \label{cube}
-\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1}, 1\leqslant i\leqslant N\rbrace=\{y(x): 0\leqslant x\leqslant 1\},
+\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1}, 1\leqslant i\leqslant N\rbrace=\{y(x): 0\leqslant x\leqslant 1\}.
 \end{equation}
-
+С использованием отображения $y(x)$ решение исходной задачи сводится к минимизации одномерной функции на отрезке $[0,1]$, т.е.
 \begin{equation}
@@
 
- У данного способа редукции размерности имеется важное свойство, необходимое для последующих расчетов: ограниченность относительных разностей функции сохраняется, т.е.
-если в области $D$ функция $\varphi(y)$ удовлетворяла условию Липшица, то на интервале [0,1] функция $\varphi(y(x))$ будет удовлетворять равномерному условию Гельдера
-
+Указанный способ редукции размерности обладает следующим важным свойством: сохраняется ограниченность относительных разностей функции , т.е.
+если в области $D$ функция $\varphi(y)$ удовлетворяла условию Липшица, то на интервале $[0,1]$ функция $\varphi(y(x))$ будет удовлетворять равномерному условию Гельдера
 \begin{equation}
 \label{holder}
-|\phi(y(x_1))-\phi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}}, 
+\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leqslant H{\left|x_1-x_2\right|}^{\frac{1}{N}}, 
  x_1,x_2\in[0,1],
 \end{equation}
-
 \begin{equation}
@@
 
- Пользуясь этим свойством, можно трактовать исходную задачу, как минимизацию одномерной гельдеровой функции $\ \ f\left(x\right)=\varphi\left(y\left(x\right)\right),\ \ x\ \in[0,\ 1]$.
+Пользуясь этим свойством, можно трактовать исходную задачу (\ref{task}) как задачу минимизации одномерной функции $f\left(x\right)=\varphi\left(y\left(x\right)\right)$, $x\in[0,1]$, удовлетворяющей условию Гельдера.
+%следующий абзац надо поправить
 Факт того, что функция задана как «черный ящик», сокращает число подходящих алгоритмов. 
@@
 
-В процессе своей работы алгоритм порождает последовательность точек $x^k$, в каждой из которых проходит вычисление значения минимизируемой функции $z^k=f(x^k)$.
-Далее под испытанием мы будем понимать подразумевать процесс вычисления значения одномерной функции, в который, помимо прочего, включается построение образа $y^k=y(x^k)$, а результатом испытания назовем пару $(x^k,\ z^k)$.
-Процесс распараллеливания построен таким образом, что при выполнении одной итерации метода одновременно проходят $P$ испытаний, $P\ \geq1$.
-Введем обозначение $k(n)$, как общее число испытаний, которые были проведены после n выполненных параллельных итераций.
+В процессе своей работы алгоритм порождает последовательность точек $\{x^k\}$, в каждой из которых проходит вычисление значений минимизируемой функции $f(x^k)$.
+В дальнейшем будем называть процесс вычисления значения одномерной функции в точке $x^k$ \textit{поисковым испытанием}.
+Испытание включает в себя также построение образа $y^k=y(x^k)$, а результатом испытания является пара $(x^k,\ z^k=f(x^k))$.
+Процесс распараллеливания построен таким образом, что при выполнении одной итерации метода одновременно проходят $P\geq1$ испытаний.
+Обозначим $k(n)$ общее число испытаний, которые были проведены после выполнения $n$ параллельных итераций.
+%текст надо исправить
 Приведем последовательность этапов параллельного алгоритма глобального поиска с модифицированием для решения задач с функциями, удовлетворяющими условию Гельдера:

### 2022-10-11 2022/NumMethOneAPI/NMP_authorGuidelinesRu_2022.tex
check Section 2

@@
 
-В процессе своей работы алгоритм порождает последовательность точек $\{x^k\}$, в каждой из которых проходит вычисление значений минимизируемой функции $f(x^k)$.
+Рассматриваемый в статье параллельный алгоритм в процессе своей работы порождает последовательность точек $\{x^k\}$, в каждой из которых проходит вычисление значений минимизируемой функции $f(x^k)$.
 В дальнейшем будем называть процесс вычисления значения одномерной функции в точке $x^k$ \textit{поисковым испытанием}.
 Испытание включает в себя также построение образа $y^k=y(x^k)$, а результатом испытания является пара $(x^k,\ z^k=f(x^k))$.
-Процесс распараллеливания построен таким образом, что при выполнении одной итерации метода одновременно проходят $P\geq1$ испытаний.
-Обозначим $k(n)$ общее число испытаний, которые были проведены после выполнения $n$ параллельных итераций.
+Процесс распараллеливания организован таким образом, что при выполнении одной итерации метода одновременно проводятся $P\geq1$ испытаний.
+Обозначим $k(n)$ общее число испытаний, которые были проведены после выполнения $n$ итераций алгоритма.
 %текст надо исправить
-Приведем последовательность этапов параллельного алгоритма глобального поиска с модифицированием для решения задач с функциями, удовлетворяющими условию Гельдера:
-Подготовительный этап.
-Изначально проводится испытание в произвольной внутренней точке $x^1\in\left[0,\ 1\right].\ $ Когда выполнено $n\geq1$ итераций, и соответствующее им число $k=k(n)$ испытаний в точках $x^i,\ 1\le\ i\le\ k$, то точки $x^{k+1},\ldots,x^{k+P}$ испытаний последующей $(n+1)$ итерации будут определять следующим образом:
 
- 1 этап. Упорядочить по координате точки уже проведенных испытаний и граничные точки:
+Вычислительная схема параллельного алгоритма для поиска глобального минимума функций, удовлетворяющих условию Гельдера, состоит в следующем.
+
+На подготовительном шаге параллельно проводятся $P$ поисковых испытаний в произвольных внутренних точках $x^1, ...,x^P$ отрезка $[0,1]$, что соответствует первой итерации  алгоритма.
+
+Если выполнено $n\geq1$ итераций, которым соответствуют $k=k(n)$ проведенных поисковых испытаний в точках $x^i, 1\leq i\leq k$, то точки $x^{k+1},\ldots,x^{k+P}$ испытаний следующей $(n+1)$-й итерации будут определять следующим образом.
+
+ Шаг 1. Перенумеровать (нижним индексом) точки ранее проведенных испытаний $x^i, 1\leq i\leq k$, а также граничные точки отрезка [0,1] в порядке возрастания координаты:
  \begin{equation}
 \label{agp1_sort}
-	0=x_0<\ x_1<\ ...\ <x_{k+1}=1,
+	0=x_0<\ x_1<\ ...\ <x_{k+1}=1.
 	\end{equation}
+	и сопоставить им значения $z_i=f(x_i)$.
 
-2 этап. Вычислить текущее значение $M$, оценивающее неизвестную константу Липшица $L$:
+Шаг 2. Вычислить текущие нижние оценки $M$ неизвестной константы Гельдера $H$:
  \begin{equation}
@@
 	\end{equation}
+где $r>1$ --- заданный параметр алгоритма.
 
-   где $r>1$ заданный параметр метода.
-
-3 этап. Найти значение характеристики $R(i)$ для всех интервалов
-
+Шаг 3. Для каждого интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить величину $R(i)$, называемую \textit{характеристикой} интервала, в соответствии с формулами
 \begin{equation}
 \label{agp3_R1}
-R(1)=2\Delta_1-4\dfrac{z_1}{M},R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
+R(1)=2\Delta_1-4\dfrac{z_1}{M}, \; R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
 \end{equation}
-
 \begin{equation}
@@
 \end{equation}
-
 где \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
 
-4 этап.  Упорядочить характеристики $R\left(i\right),\ 1\ \le\ i\ \le\ k+1$ в порядке невозрастания 
+Шаг 4.  Упорядочить характеристики $R\left(i\right),\ 1\leq i \leq k+1,$ в порядке невозрастания 
 \begin{equation}
@@
 \end{equation}	
+и выбрать $P$ интервалов с номерами $t_j,\ 1\le\ j\le\ P$, с наибольшими значениями характеристики.
 
-и выбрать P интервалов с номерами $t_j,\ 1\le\ j\le\ P$, значение характеристики в которых наибольшее.
-
-5 этап. В выбранных интервалах вычислить точки $x^{k+j},\ 1\le\ j\le\ P$ и провести в них новые испытания:
-
+Шаг 5. В выбранных интервалах вычислить точки $x^{k+j},\ 1\leq j\leq P,$ в соответствии с формулами
 \begin{equation}
@@
 \end{equation}	
-
 	\begin{equation}
@@
 
-Важно отметить, что в рассматриваемых задачах самый трудоемкий и времязатратный этап алгоритма – это вычисление значения функции в точке.
-Поэтому параллельное вычисление этих значений положительно влияет на скорость работы алгоритма.
+Очередные испытания проводятся параллельно, в вычисленных по формулам (\ref{agp5_x1}), (\ref{agp4_xi}) точках $x^{k+j},\ 1\leq j\leq P$.
+Отметим, что в прикладных оптимизационных задачах процесс проведения испытания является, как правило, значительно более трудоемким по сравнению с работой вычислительных правил алгоритма, поэтому параллельное проведение испытаний ускоряет работу алгоритма.
 
-Используемые критерии остановки:
+Алгоритм останавливает свою работу в случае, если хотя бы для одного значения $t_j,\ 1\le\ j\le\ P$ из (\ref{agp4_R_sort}) выполняется условие \(\Delta_{t_j} < \varepsilon\).
+Данный критерий остановки (наряду с обычным для итерационных методов критерием, ограничивающим число выполненных итераций) используется в задачах оптимизации, в которых точка глобального минимума $x^*$ заранее неизвестна.
 
-	по длине интервала, как только происходит выполнение условия \(\Delta_{t_j}\leqslant \varepsilon\) хотя бы для какого-то номера $t_j,\ 1\le\ j\le\ P$.
-Основной критерий остановки, используется в прикладных задачах оптимизации и в тестовых наборах, для которых точка глобального минимума заранее неизвестна.
+При решении тестовых задач, в которых точка глобального минимума $x^*$ является  известной, можно использовать также и критерий остановки по попаданию в окрестность глобального минимума.
+В этом случае метод прекращает работу, если если хотя бы для одного значения $t_j,\ 1\le\ j\le\ P$ из (\ref{agp4_R_sort}) выполняется условие $\left|x_{t_j}-\ x^\ast\right| < \varepsilon.$
 
-	по попаданию в окрестность глобального минимума, $\left|x_{t_j}-\ x^\ast\right|<\varepsilon,\ \ t_j,\ 1\le\ j\le\ P$, где $x^\ast$ -- точка глобального минимума, если ее значение известно.
-Используется в тестовых задачах, когда $x^\ast$ известна до начала работы алгоритма.
-
-Как оценка глобально-оптимального решения рассматриваемой задачи (1) выбираются значения
-
+В качестве итоговой оценки глобально-оптимального решения рассматриваемой задачи выбираются значения 
 \begin{equation}
-f_k^*=\min_{1\leqslant i \leqslant k}f(x_i), x_k^*=arg \min_{1\leqslant i \leqslant k}f(x_i).
+f_k^*=\min_{1\leqslant i \leqslant k}f(x_i), \; x_k^*=arg \min_{1\leqslant i \leqslant k}f(x_i).
 \end{equation}
@@
 \section{Теоретические оценки ускорения параллельного алгоритма}
-Опишем (по книге [\ref{rfa:rulit:lit2}]) теоретические свойства параллельного алгоритма.
+
+Опишем (в соответствии с [\ref{rfa:rulit:lit2}]) теоретические свойства рассматриваемого параллельного алгоритма.
 В решаемых задачах время выполнения одного испытания существенно превышает время обработки ее результата.
@@
 Точки других испытаний, как правило, могут не совпадать с генерируемыми последовательным алгоритмом.
-Поэтому будем называть такие испытания излишними, а количество
+Поэтому будем называть такие испытания \textit{избыточными}, а величину
 \begin{displaymath}
@@
 \end{displaymath}
-будем называть избыточностью метода.
+будем называть \textit{избыточностью} метода.
 
-Следующие утверждения из [\ref{rfa:rulit:lit2}] определяют степень распараллеливания $p$, которая соответствует неизбыточному (т.е. с нулевой избыточностью) распараллеливанию.
+Следующие утверждения из [\ref{rfa:rulit:lit2}] определяют степень распараллеливания $p$, которая соответствует безызбыточному (т.е. с нулевой избыточностью) распараллеливанию.
 
@@
 
-\textbf{Теорема}. Пусть $x^*$ — точка глобального минимума, а $x^{\prime}$ — точка локального минимума функции $f(x)$, и пусть выполнены следующие условия:
+\textbf{Теорема}. Пусть $x^*$ --- точка глобального минимума, а $x^{\prime}$ --- точка локального минимума функции $f(x)$, и пусть выполнены следующие условия:
     \begin{enumerate}
@@
 
-Тогда параллельный алгоритм при $p=2$ будет неизбыточным (т.е. $s(2)=2$, $\lambda(2)=0$) до тех пор, пока выполняется условие
-
+Тогда параллельный алгоритм при $p=2$ будет безызбыточным (т.е. $s(2)=2$, $\lambda(2)=0$), пока выполняется условие
 \begin{equation} \label{p_two_ref}

### 2022-10-12 2022/NumMethOneAPI/NMP_authorGuidelinesRu_2022.tex
Fix the last section and reference list

@@
 }%Обновить перевод названия
-    {Solving time-consuming multidimensional global optimization problems with Intel OneApi tools}
+    {On implementation of the parallel global optimization algorithm with the Intel oneAPI toolkit}
 
@@
 %обновить перевод - сейчас он для старой аннотации
-    {The paper considers the process of solving a series of time-consuming global optimization problems.
-We suppose that the objective function satisfies the Lipschitz condition with a priory unknown Lipschitz constant.
-Intel OneApi tools, that allows one to write the same code for both the central processor and the graphics accelerator, are chosen for implementing the parallel global optimization algorithm.
-To solve multidimensional problems, we use an approach based on the idea of dimensionality reduction using the Peano curve, which continuously maps an interval of the real axis onto an n-dimensional cube.}
+    {The paper considers the parallel global optimization algorithm ans discusses its implementation with the Intel oneAPI toolkit.
+We suppose that the objective function is given as a black-box and satisfies the Lipschitz condition.
+The parallel algorithm presented in the paper uses the scheme of dimensionality reduction employing the Peano curve, which continuously maps an interval of the real axis onto a hypercube.
+The Intel OneApi tools, that allows one to write the same code for both the central processor and the graphics accelerator, were used for implementation of the parallel global optimization algorithm.
+The results of numerical experiments obtained by solving a series of time-consuming multiextremal optimization problems are presented.}
 
@@
 \Citation{Баркалов К.А., Лебедев И.Г., Силенко Я.В.
-    Решение трудоемких задач многомерной глобальной оптимизации с использованием набора инструментов Intel oneAPI~//
+    О реализации параллельного алгоритма глобальной оптимизации с использованием набора инструментов Intel oneAPI~//
     Вычислительные методы и программирование. 2022.
@@
     {K.~A.~Barkalov, I.~G.~Lebedev, Ya.~V.~Silenko,
-    Solving time-consuming multidimensional global optimization problems with Intel OneApi tools 
-    for the journal ``Numerical methods and programming,''
+    On implementation of the parallel global optimization algorithm with the Intel oneAPI toolkit
+		``Numerical methods and programming,''
     Numerical Methods and Programming. \textbf{??} (?), 
@@
 
-\UDC{<указывает автор>}
+\UDC{519.853.4}
 \DOI{10.26089/NumMet.v??r???}% заполняет редакция
@@
 \YEAR{2022}
-\Received{16 июня 2022 г.}{June 16, 2022}
-\Accepted{13 октября 2022 г.}{October 13, 2022}
+\Received{16 октября 2022 г.}{October 16, 2022}
+\Accepted{?? ??????? 2022 г.}{??????? ??, 2022}
 
@@
 \Institution{Нижегородский государственный университет им. Н.И. Лобачевского}
-{Nizhny Novgorod State University N.I. Lobachevsky}
+{Lobachevsky State University of Nizhny Novgorod}
 \Address{пр. Гагарина, д. 23.}{Gagarin Ave., 23.}
@@
 \AcademicDegree{
-Доктор технических наук}{Ph.D.}
-\Position{доцент}{docent}
+Доктор технических наук}{D.Sc.}
+\Position{профессор}{professor}
 \Orcid{0000-0001-5273-2471}
-\Email{konstantinbarkalov@yandex.ru}
+\Email{konstantin.barkalov@itmm.unn.ru}
 
@@
 
-Задачи поиска экстремумов многоэкстремальных функций часто возникают в различных областях науки, таких как химия, физика, машиностроение и т.д.
+Задачи поиска минимума многоэкстремальных функций часто возникают в различных областях науки, таких как химия, физика и т.д.
+В частности, такие задачи возникают при разработке новых лекарственных препаратов [\ref{rfa:rulit:litSulimov}], при поиске конфигураций новых химических соединений [\ref{rfa:rulit:litPosypkin1},\ref{rfa:rulit:litPosypkin2}].
+Задачи глобальной оптимизации возникают и во многих других приложениях [\ref{rfa:rulit:lit4}].
+Традиционной сферой применения методов глобальной оптимизации стала идентификация параметров математических моделей по данным экспериментов.
+В задачах такого вида требуется провести поиск значений неизвестных параметров модели, при которых результаты расчетов близки к результатам, полученным экспериментально [\ref{rfa:rulit:litUfa}].
 
-В рамках данной работы рассматривается задача поиска глобального минимума $y^*$ функции $\varphi(y)$ в гиперинтервале $D$, т.е. 
+Мы будем рассматривать задачу глобальной оптимизации в следующей постановке: требуется найти глобальный минимума$y^*$ функции $\varphi(y)$ в гиперинтервале $D$, т.е. 
 \begin{equation}
@@
 
-%комментарий про разные способы редукции размерности - ссылки на Сергеева, Жилинскаса
-
-Сведение многомерной задачи (\ref{task}) к одномерной проводится за счет использования редукции размерности при помощи кривой Пеано [\ref{rfa:rulit:lit1}, \ref{rfa:rulit:lit2}] $y(x)$, которая непрерывно и однозначно отображает отрезок вещественной оси $[0,1]$ на $N$-мерный гиперкуб:
+Многие известные алгоритмы липшицевой глобальной оптимизации основаны на идее редукции размерности и адаптации одномерных алгоритмов для решения многомерных задач [\ref{rfa:rulit:lit16},\ref{rfa:rulit:lit10}].
+В данной работе мы будем использовать подход из [\ref{rfa:rulit:lit2}, \ref{rfa:rulit:lit1}], основанный на идее редукции размерности с помощью кривой Пеано  
+$y(x)$, которая непрерывно и однозначно отображает отрезок вещественной оси $[0,1]$ на $N$-мерный гиперкуб:
 \begin{equation}
@@
 \end{equation}
-С использованием отображения $y(x)$ решение исходной задачи сводится к минимизации одномерной функции на отрезке $[0,1]$, т.е.
+С использованием отображения $y(x)$ решение исходной задачи (\ref{task}) сводится к минимизации одномерной функции на отрезке $[0,1]$, т.е.
 \begin{equation}
@@
 Пользуясь этим свойством, можно трактовать исходную задачу (\ref{task}) как задачу минимизации одномерной функции $f\left(x\right)=\varphi\left(y\left(x\right)\right)$, $x\in[0,1]$, удовлетворяющей условию Гельдера.
-%следующий абзац надо поправить
-Факт того, что функция задана как «черный ящик», сокращает число подходящих алгоритмов. 
-Для исследования выбран алгоритм глобального поиска [\ref{rfa:rulit:lit3}, \ref{rfa:rulit:lit4}], который является одним из эффективных методов глобальной оптимизации, т.к.
-при решении многих задач опережает (по числу итераций) другие методы аналогичного назначения [\ref{rfa:rulit:lit5} – \ref{rfa:rulit:lit9}].
-Алгоритм глобального поиска относится к классу характеристических алгоритмов оптимизации, хорошо масштабируется и отлично подходит для работы на параллельных вычислительных кластерах.
+Для решения указанной задачи минимизации может быть применен \textit{алгоритм глобального поиска} [\ref{rfa:rulit:lit3}], который является эффективным алгоритмом, т.к.
+при решении сложных многоэкстремальных задач опережает (по числу итераций, требующихся для корректного решения задачи с фиксированной точностью) другие методы аналогичного назначения [\ref{rfa:rulit:lit5}].
+Алгоритм глобального поиска является весьма гибким и отлично подходит для реализации на параллельных вычислительных системах.
 
@@
 
-Рассматриваемый в статье параллельный алгоритм в процессе своей работы порождает последовательность точек $\{x^k\}$, в каждой из которых проходит вычисление значений минимизируемой функции $f(x^k)$.
+Рассматриваемый в статье параллельный алгоритм глобального поиска в процессе своей работы порождает последовательность точек $\{x^k\}$, в каждой из которых проходит вычисление значений минимизируемой функции $f(x^k)$.
 В дальнейшем будем называть процесс вычисления значения одномерной функции в точке $x^k$ \textit{поисковым испытанием}.
-Испытание включает в себя также построение образа $y^k=y(x^k)$, а результатом испытания является пара $(x^k,\ z^k=f(x^k))$.
-Процесс распараллеливания организован таким образом, что при выполнении одной итерации метода одновременно проводятся $P\geq1$ испытаний.
-Обозначим $k(n)$ общее число испытаний, которые были проведены после выполнения $n$ итераций алгоритма.
-%текст надо исправить
+Испытание включает в себя также построение образа $y^k=y(x^k)$, а результатом испытания является пара $(x^k,\ z^k)$, где $z^k=\varphi(y(x^k))$.
+Процесс распараллеливания организован таким образом, что при выполнении одной итерации метода одновременно проводятся $p\geq1$ испытаний.
 
-Вычислительная схема параллельного алгоритма для поиска глобального минимума функций, удовлетворяющих условию Гельдера, состоит в следующем.
+Для описания вычислительной схемы параллельного алгоритма глобального поиска обозначим $k(n)$ общее число испытаний, которые были проведены после выполнения $n$ итераций алгоритма.
 
-На подготовительном шаге параллельно проводятся $P$ поисковых испытаний в произвольных внутренних точках $x^1, ...,x^P$ отрезка $[0,1]$, что соответствует первой итерации  алгоритма.
+На подготовительном шаге параллельно проводятся $p$ поисковых испытаний в произвольных внутренних точках $x^1, ...,x^p$ отрезка $[0,1]$, что соответствует первой итерации  алгоритма.
 
-Если выполнено $n\geq1$ итераций, которым соответствуют $k=k(n)$ проведенных поисковых испытаний в точках $x^i, 1\leq i\leq k$, то точки $x^{k+1},\ldots,x^{k+P}$ испытаний следующей $(n+1)$-й итерации будут определять следующим образом.
+Если выполнено $n\geq1$ итераций, которым соответствуют $k=k(n)$ проведенных поисковых испытаний в точках $x^i, 1\leq i\leq k$, то точки $x^{k+1},\ldots,x^{k+p}$ испытаний следующей $(n+1)$-й итерации будут определять следующим образом.
 
@@
 \end{equation}	
-и выбрать $P$ интервалов с номерами $t_j,\ 1\le\ j\le\ P$, с наибольшими значениями характеристики.
+и выбрать $p$ интервалов с номерами $t_j,\ 1\le\ j\le\ p$, с наибольшими значениями характеристики.
 
-Шаг 5. В выбранных интервалах вычислить точки $x^{k+j},\ 1\leq j\leq P,$ в соответствии с формулами
+Шаг 5. В выбранных интервалах вычислить точки $x^{k+j},\ 1\leq j\leq p,$ в соответствии с формулами
 \begin{equation}
@@
 
-Очередные испытания проводятся параллельно, в вычисленных по формулам (\ref{agp5_x1}), (\ref{agp4_xi}) точках $x^{k+j},\ 1\leq j\leq P$.
+Очередные испытания проводятся параллельно, в вычисленных по формулам (\ref{agp5_x1}), (\ref{agp4_xi}) точках $x^{k+j},\ 1\leq j\leq p$.
 Отметим, что в прикладных оптимизационных задачах процесс проведения испытания является, как правило, значительно более трудоемким по сравнению с работой вычислительных правил алгоритма, поэтому параллельное проведение испытаний ускоряет работу алгоритма.
 
-Алгоритм останавливает свою работу в случае, если хотя бы для одного значения $t_j,\ 1\le\ j\le\ P$ из (\ref{agp4_R_sort}) выполняется условие \(\Delta_{t_j} < \varepsilon\).
+Алгоритм останавливает свою работу в случае, если хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}) выполняется условие \(\Delta_{t_j} < \varepsilon\).
 Данный критерий остановки (наряду с обычным для итерационных методов критерием, ограничивающим число выполненных итераций) используется в задачах оптимизации, в которых точка глобального минимума $x^*$ заранее неизвестна.
@@
 При решении тестовых задач, в которых точка глобального минимума $x^*$ является  известной, можно использовать также и критерий остановки по попаданию в окрестность глобального минимума.
-В этом случае метод прекращает работу, если если хотя бы для одного значения $t_j,\ 1\le\ j\le\ P$ из (\ref{agp4_R_sort}) выполняется условие $\left|x_{t_j}-\ x^\ast\right| < \varepsilon.$
+В этом случае метод прекращает работу, если если хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}) выполняется условие $\left|x_{t_j}-\ x^\ast\right| < \varepsilon.$
 
@@
 Опишем (в соответствии с [\ref{rfa:rulit:lit2}]) теоретические свойства рассматриваемого параллельного алгоритма.
-В решаемых задачах время выполнения одного испытания существенно превышает время обработки ее результата.
-Таким образом, ускорение по испытаниям будет:
+В решаемых задачах время выполнения одного испытания существенно превышает время обработки его результата.
+Поэтому ускорение параллельного алгоритма можно определить как \textit{ускорение по испытаниям} $s(p)$,
+\begin{equation} \label{par_trl_ref}
+s(p) = \frac{n(1) \cdot p}{n(p)},
+\end{equation}
+где $n(1)$ --- количество испытаний, выполненных последовательным алгоритмом, а $n(p)$ --- количество испытаний, выполненных параллельным алгоритмом с использованием $p$ параллельных вычислительных устройств.
+Ускорение по испытаниям $s(p)$ является ключевой характеристикой эффективности параллельного алгоритма глобального поиска.
 
-\begin{equation} \label{par_trl_ref}
-s(p) = \frac{n(1) \cdot p}{n(p)}
-\end{equation}
-
-$s(p)$ является ключевой характеристикой эффективности параллельного алгоритма глобального поиска.
-Где $n(1)$ — количество испытаний, выполненных последовательным алгоритмом, а $n(p)$ — количество испытаний, выполненных параллельным алгоритмом с использованием $p$ параллельных устройств.
-
-Очевидно, что количество испытаний $n(p)$ для алгоритмов с разной степенью распараллеливания $p$ будут отличаться друг от друга.
-Действительно, последовательный алгоритм имеет полную информацию, полученную на предыдущих $k$ итерациях, при выборе точки $x^{k+1}$ следующего $(k+1)-ого$ испытания.
-Параллельный алгоритм выбирает не одну, а $p$ точек $x^{k+j}$, $1 \leq j \leq p,$ на основе той же информации.
-Это означает, что выбор точек $x^{k+j}$ производится без информации о результатах испытаний в точках $x^{k+i}$, $1 \leq i < j$.
+Отметим, что количество испытаний $n(p)$ для алгоритмов с разной степенью распараллеливания $p$ будут отличаться друг от друга.
+Действительно, последовательный алгоритм при выборе точки $x^{k+1}$ следующего $(k+1)$-го испытания имеет полную информацию, полученную на предыдущих $k$ итерациях.
+Параллельный алгоритм выбирает не одну, а $p$ точек $x^{k+1},..., x^{k+p}$  на основе той же информации.
+Это означает, что выбор точек $x^{k+j}, \; 1 < j \leq p,$ производится без информации о результатах испытаний в точках $x^{k+i},\; 1 \leq i < j$.
 Только первая точка $x^{k+1}$ будет соответствовать точке, выбранной последовательным алгоритмом.
-Точки других испытаний, как правило, могут не совпадать с генерируемыми последовательным алгоритмом.
+Точки других испытаний могут не совпадать с точками, генерируемыми последовательным алгоритмом.
 Поэтому будем называть такие испытания \textit{избыточными}, а величину
@@
 
-Следующие утверждения из [\ref{rfa:rulit:lit2}] определяют степень распараллеливания $p$, которая соответствует безызбыточному (т.е. с нулевой избыточностью) распараллеливанию.
+Следующее утверждение определяет степень распараллеливания $p$, которая соответствует безызбыточному (т.е.
+с нулевой избыточностью) распараллеливанию (доказательство утверждения см.
+в [\ref{rfa:rulit:lit2}]).
 
-Обозначим серии испытаний, генерируемые последовательным алгоритмом и параллельным при решении одной и той же задачи с $\varepsilon=0$ в условиях остановки, как $\{x^k\}$ и $\{y^m \}$ соответственно.
+Обозначим испытания, генерируемые последовательным и параллельным алгоритмами при решении одной и той же задачи с $\varepsilon=0$ в условиях остановки, как $\{x^k\}$ и $\{y^m \}$ соответственно.
 
@@
             \end{equation}
-        где $H$ — постоянная Гёльдера целевой одномерной функции.
+        где $H$ — постоянная Гельдера для одномерной целевой функции.
     \end{enumerate}
-
 Тогда параллельный алгоритм при $p=2$ будет безызбыточным (т.е. $s(2)=2$, $\lambda(2)=0$), пока выполняется условие
@@
 \end{equation}
-
-где $t_j$ определяется в соответствии с шагом 4 алгоритма.
+где $t_j$ определяется в соответствии с (\ref{agp4_R_sort}).
 
 \textbf{Следствие}.
-
 Пусть целевая функция $f(x)$ имеет $Q$ локальных минимумов $\{x_1^{\prime},...,x_Q^{\prime}\}$, для которых выполняется условие \eqref{first_s_ref} и пусть существуют точки испытания $y^{n_i}$, $1 \leq i \leq Q,$ такие, что
@@
 \end{gather}
-Тогда при выполнении условий теоремы, параллельный алгоритм со степенью распараллеливания $Q+1$ будет неизбыточным (т.е.
-$s(Q+1)=Q+1, \; \lambda(Q+1) =0$) до тех пор, пока не будет выполнено условие \eqref{p_two_ref}.
+Тогда при выполнении условий теоремы параллельный алгоритм со степенью распараллеливания $Q+1$ будет безызбыточным (т.е.
+$s(Q+1)=Q+1, \; \lambda(Q+1) =0$) до тех пор, пока выполняется условие \eqref{p_two_ref}.
 
 Следствие из теоремы играет особую роль при решении многомерных задач, сводящихся к одномерным с помощью отображения $y(x)$.
-Отображение $y(x)$, являясь аппроксимацией кривой Пеано, имеет эффект "расщепления" точки глобального минимума $y^* \in D$ на несколько (до $2^N$) прообразов в интервале [0,1].
-Применяя параллельный алгоритм для минимизации редуцированной одномерной задачи, можно ожидать нулевую избыточность при степени распараллеливания до $2^N+1$.
+Отображение $y(x)$, являясь аппроксимацией кривой Пеано, характеризуется эффектом <<расщепления>> точки глобального минимума $y^* \in D$ на несколько (до $2^N$) прообразов на отрезке $[0,1]$.
+Применяя параллельный алгоритм для минимизации редуцированной одномерной функции, можно ожидать нулевую избыточность при степени распараллеливания до $2^N+1$.
 
@@
 Рассматриваемые многомерные многоэкстремальные задачи обладают высокой трудоемкостью численного решения, поскольку при увеличении размерности задачи наблюдается экспоненциальный рост вычислительных затрат.
-Вместе с тем быстрый процесс развития современных вычислительных средств, включая распараллеливание, предоставляет всё больше различных новых возможностей для решения проблем оптимизации.
-Однако на фоне этого возникает задача эффективного распараллеливания, а многообразие различных типов современных ускорителей и средств разработки для них дают пользователям немалый выбор [\ref{rfa:rulit:lit10}].
+Вместе с тем быстрый прогресс параллельных вычислительных средств предоставляет всё больше различных новых возможностей для решения сложных задач.
+Однако на фоне этого возникает вопрос эффективного распараллеливания, а многообразие различных типов устройств и средств разработки ставят пользователя перед проблемой выбора одной из них [\ref{rfa:rulit:lit10}].
 
-Но перейдя от частной задачи к общей ситуации, разберем, какие предлагаются варианты решения.
-Для обеспечения высокой производительности вычислений в рамках каких-то новых рабочих задач требуются различные вычислительные архитектуры.
+В общем случае, для обеспечения высокой производительности вычислений в рамках решения новых задач требуются различные вычислительные архитектуры.
 Например, компанией Intel предлагается следующая классификация ускорителей: скалярные (CPU), векторные (GPU), матричные и ПЛИС (FPGA).
@@
 Таким образом, первые этапы рассматриваемого алгоритма выполняются на центральном процессоре (CPU).
-Далее, полученные в ходе выполнения одной итерации новые P координат из интервалов с наибольшими характеристиками, передаются с помощью промежуточного буфера на устройство (CPU, GPU, FPGA), выбранное для исполнения распараллеленного блока кода, для вычисления значения функции в них.
+Далее, полученные в ходе выполнения одной итерации новые $p$ координат из интервалов с наибольшими характеристиками, передаются с помощью промежуточного буфера на устройство (CPU, GPU, FPGA), выбранное для исполнения распараллеленного блока кода, для вычисления значения функции в них.
 Затем найденные значения функции в этих точках передаются через промежуточный буфер обратно на центральный процессор.
@@
 
-В рамках работы были проведены вычислительные эксперименты на персональном компьютере с процессором Intel Core i5-10600 3.3GHz, 32 GB оперативной памяти и встроенной графической картой Intel UHD Graphics 630.
+Вычислительные эксперименты были проведены на компьютере с процессором Intel Core i5-10600 3.3GHz, 32 GB оперативной памяти и графической картой Intel UHD Graphics 630.
 Для получения исполняемого программного кода использовался компилятор Intel oneAPI DPC++ 2021.1.
-Вычислительные эксперименты выполнялись с использованием разработанного программного комплекса Globalizer [\ref{rfa:rulit:lit14}, \ref{rfa:rulit:lit15}].
+Вычислительные эксперименты выполнялись с использованием  программной системы Globalizer [\ref{rfa:rulit:lit14}].
 
-Большинство известных тестовых задач из области многомерной глобальной оптимизации характеризуются небольшим временем вычисления значений целевой функции.
-Из-за этого при проведении экспериментов бывает сложно понять с чем может быть связано относительно большое время выполнения параллельного алгоритма: с появившимися, сравнительно с последовательной версией, накладными расходами или причина в неэффективном распараллеливании.
-В реальных же задачах вычисление функции самая трудоемкая и время затратная операция, поэтому такой проблемы не возникает.
-
-Нами предлагается вычислительно трудоемкая модификация существующих тестовых задач, расчет целевой функции заключается в интегрировании исходной тестовой функции по части параметров.
-Для этого изначально порождается задача, размерность которой в два раза больше искомой.
-При вычислении значения функции, первые $N$ координат фиксируются, а по остальным производится численное интегрирование по области определения функции.
+Большинство известных тестовых задач из области многомерной глобальной оптимизации характеризуются небольшим временем вычисления значений целевой функции, которое сопоставимо со временем работы расчетных правил алгоритма.
+%Из-за этого при проведении экспериментов бывает сложно понять, с чем может быть связано относительно большое время выполнения параллельного алгоритма: с появившимися, сравнительно с последовательной версией, накладными расходами или причина в неэффективном распараллеливании.
+В прикладных оптимизационных задачах вычисление функции является трудоемкой операцией, здесь время работы алгоритма будет значительное меньше времени проведения одного испытания. 
+В связи с этим нами была предложена модификация существующих тестовых задач, в которых требуется провести интегрирование исходной тестовой функции по части параметров.
+Для этого изначально порождается функция $\psi(y), y\in R^{2N},$ размерности $2N$, в которой первые $N$ координат фиксируются, а по остальным производится численное интегрирование по области определения функции.
 Для интегрирования используется метод средних прямоугольников.
@@
 \begin{displaymath}
-\underset{{{i}_{1}}=0}{\overset{M-1}{\mathop \sum }}\,\underset{{{i}_{2=0}}}{\overset{M-1}{\mathop \sum }}\,..\underset{{{i}_{N}}=0}{\overset{M-1}{\mathop \sum }}\,\left( \left( \underset{j=1}{\overset{2N}{\mathop \prod }}\,{{h}_{j}} \right)\psi \left( {{y}_{1}},{{y}_{2}},~..,{{y}_{N}},~{{y}_{{{\left( N+1 \right)}_{{{i}_{1}}}}}},{{y}_{{{\left( N+2 \right)}_{{{i}_{2}}}}}}~,~..,{{y}_{{{\left( 2N \right)}_{{{i}_{N}}}}}} \right) \right),
+\underset{{{i}_{1}}=0}{\overset{M-1}{\mathop \sum }}\,\underset{{{i}_{2=0}}}{\overset{M-1}{\mathop \sum }}\,...\underset{{{i}_{N}}=0}{\overset{M-1}{\mathop \sum }}\,\left( \left( \underset{j=1}{\overset{2N}{\mathop \prod }}\,{{h}_{j}} \right)\psi \left( {{y}_{1}},{{y}_{2}},~..,{{y}_{N}},~{{y}_{{{\left( N+1 \right)}_{{{i}_{1}}}}}},{{y}_{{{\left( N+2 \right)}_{{{i}_{2}}}}}}~,~..,{{y}_{{{\left( 2N \right)}_{{{i}_{N}}}}}} \right) \right),
 \end{displaymath}
 \begin{displaymath}
-D=\left\{ y~\in ~{{R}^{2N}}:~{{a}_{i}}\le ~~{{y}_{i}}\le ~~{{b}_{i}},~1\le i\le 2N \right\},
+D=\left\{ y~\in ~{{R}^{2N}}:~{{a}_{i}}\le {{y}_{i}}\le {{b}_{i}},~1\le i\le 2N \right\},
 \end{displaymath}
 \begin{equation}
-{{y}_{{{\left( N+k \right)}_{{{i}_{k}}}}}}={{a}_{k}}+{{i}_{k}}*{{h}_{k}}+\frac{{{h}_{k}}}{2},
+{{y}_{{{\left( N+k \right)}_{{{i}_{k}}}}}}={{a}_{k}}+{{i}_{k}}{{h}_{k}}+\frac{{{h}_{k}}}{2},
 \end{equation}
 \begin{equation}
-{{h}_{k}}=~\frac{{{b}_{k}}-~{{a}_{k}}}{M},
+{{h}_{k}}=~\frac{{{b}_{k}}-~{{a}_{k}}}{M}.
 \end{equation}
-
-где $M$ -- количество участков интегрирования по одной координате, а $\psi$ – исходная тестовая функция.
+Здесь $M$ --- количество участков интегрирования по одной координате, а $\psi(y)$ --- исходная тестовая функция.
 Очевидно, чем больше значение $M$, тем больше проводится вычислений.
@@
 
-Вначале проведем вычислительные эксперименты на классической задаче – функции Растригина. Она задается простой формулой, сепарабельна и не имеет ограничений по размерности.
-
+Вначале проведем вычислительные эксперименты с использованием классической тестовой функции Растригина
 \begin{equation}
@@
 \end{equation}
-
 \begin{displaymath}
@@
 \end{displaymath}
-
 На рис.
-\ref{fig:s2} изображены линии уровня двухмерной функции Растригина (слева) и интегрированной четырехмерной функции Растригина (справа).
-Как можно видеть, новая функция не сильно отличается от двухмерного оригинала, и даже глобальный минимум остался прежним – точка (0, 0).
+\ref{fig:s2} изображены линии уровня двухмерной функции Растригина (слева) и модифицированной функции Растригина, в которой проводилось интегрирование по двум переменным (справа).
+Как можно видеть, новая функция не сильно отличается от исходной --- ее глобальный минимум остался прежним, в точке (0,0).
 
@@
   \includegraphics[width=1.0\linewidth]{./pic/s2.png}
-  \caption{Линии уровня двумерной функции Растригина (слева) и интегрированной четырехмерной функции Растригина (справа).}
+  \caption{Линии уровня двумерной функции Растригина (слева) и модифицированной четырехмерной функции Растригина (справа).}
   \label{fig:s2}  
@@
 
-Число итераций параллельного алгоритма глобального поиска (ПАГП) было ограничено 1000000, точность поиска $\varepsilon=0.01$ и параметр метода $r\ =\ 3$.
-Размерность задачи $N = 4$ и $5$.
-При вычислениях на CPU число потоков варьировалось от 1 до 8, а при вычислении на GPU от 256 до 1024.
-Использовался критерий остановки по попаданию в окрестность.
-В таблице \ref{table1} приведено число итераций ПАГП, в таблице \ref{table2} – ускорение по сравнению с однопоточным запуском.
+При проведении экспериментов число испытаний параллельного алгоритма глобального поиска (ПАГП), было ограничено величиной $K_{max}=10^6$, точность поиска установлена $\varepsilon=0.01$, использовался параметр метода $r\ =\ 3$.
+Размерность решаемых задач была $N = 4$ и $N = 5$.
+При вычислениях на CPU число потоков варьировалось от 1 до 8, а при вычислении на GPU --- от 256 до 1024.
+Так как решались тестовые задачи, то использовался критерий остановки по попаданию в окрестность известного решения.
+В таблице \ref{table1} приведено число итераций ПАГП, в таблице \ref{table2} --- его ускорение по сравнению с однопоточным запуском.
 
@@
 
-Как можно видеть из полученных результатов, использование инструментов Intel oneAPI, при распараллеливании алгоритма глобального поиска, показала хорошие результаты на тестовой функции.
-
-Далее проведем эксперименты на серии задач полученных интегрированием функций из генератора GKLS.
-Данный генератор описан в работах [\ref{rfa:rulit:lit16}, \ref{rfa:rulit:lit17}], дает возможность создавать серии задач многоэкстремальной оптимизации и заранее задавать их свойства, такие как: размерность задачи, количество локальных минимумов, размеры их областей притяжения, координата точки глобального минимума, значение функции в ней и т.п.
+Далее были проведены эксперименты на серии задач, полученных интегрированием функций из генератора GKLS.
+Данный генератор описан в [\ref{rfa:rulit:lit16}].
+С его помощью можно порождать задачи многоэкстремальной оптимизации и варьировать их свойства: размерность, количество локальных минимумов, размеры их областей притяжения, координаты точки глобального минимума, значение функции в ней и т.д.
 На рис.
-\ref{fig:s3} изображены линии уровня двухмерной функции GKLS номер 9 (слева) и интегрированной функции GKLS номер 9, полученной интегрированием четырехмерной функции по последним двум параметрам (справа).
+\ref{fig:s3} изображены линии уровня двухмерной функции GKLS (слева) и модифицированной функции GKLS, полученной интегрированием четырехмерной функции по последним двум параметрам (справа).
 
@@
   \includegraphics[width=1.0\linewidth]{./pic/s3.png}
-  \caption{Функции GKLS (слева) и интегрированной функции GKLS (справа).}
+  \caption{Функции GKLS (слева) и модифицированной функции GKLS (справа).}
   \label{fig:s3}  
@@
 
-Число итераций ПАГП было ограничено 100000, точность поиска $\varepsilon=0.01$ и параметр метода $r\ =\ 4$.
-Размерность задачи $N = 4$ и 5.
+Число итераций ПАГП было ограничено величиной $10^5$, точность поиска $\varepsilon=0.01$, параметр $r\ =\ 4$.
+Было сгенерировано две серии по 100 задач размерности $N = 4$ и $N = 5$.
 При вычислениях на CPU число потоков варьировалось от 1 до 8, а при вычислении на GPU от 256 до 1024.
@@
     \centering
-    \caption{Ускорение ПАГП при решении серии интегрированных функций GKLS.}
+    \caption{Ускорение ПАГП при решении серии модифицированных функций GKLS.}
     \label{table3}
@@
 \end{table}
+
+Полученные результаты подтверждают, что использование инструментов Intel oneAPI для реализации параллельного алгоритма глобального поиска позволяет написать одну версию кода, которая будет будет показывать хорошее ускорение как при использовании центрального процессора, так и при использовании графической карты.
 
@@
 
-\rlitem{rfa:rulit:lit1}{
-\textit{Sergeyev Ya.D., Strongin R.G, Lera D.} Introduction to global optimization exploiting space-filling curves. Springer, 2013. 125 p.}
+\rlitem{rfa:rulit:litSulimov}{
+\textit{Kutov D., Sulimov A., Sulimov V.} Supercomputer docking: investigation of low energy minima of protein-ligand complexes.
+Supercomputing Frontiers and Innovations, 2018.
+vol.
+5(3).
+P.
+134--137.}
 
-\rlitem{rfa:rulit:lit2}{
-\textit{Strongin R.G., Sergeyev Ya.D.} Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, 2000. 704 p.}
+\rlitem{rfa:rulit:litPosypkin1}{
+\textit{Абгарян К.К., Посыпкин М.
+А.} Применение оптимизационных методов для решения задач параметрической идентификации потенциалов межатомного взаимодействия.
+Ж.
+вычисл.
+матем.
+и матем.
+физ., 2014.
+Т.
+54, №12.
+С.1994--2001.}
 
-\rlitem{rfa:rulit:lit3}{
-\textit{Стронгин Р.Г., Гергель В.А., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013. 280 с.}
+\rlitem{rfa:rulit:litPosypkin2}{
+\textit{Евтушенко Ю.Г., Лурье С.А., Посыпкин М.А., Соляев Ю.О.} Применение методов оптимизации для поиска равновесных состояний двумерных кристаллов.
+Ж.
+вычисл.
+матем.
+и матем.
+физ., 2016.
+T.
+56, №12.
+С.
+2032--2041.}
 
 \rlitem{rfa:rulit:lit4}{
-\textit{Pinter (Ed.), J.D.} Global Optimization: Scientific and Engineering Case Studies. Springer, 2006. 546 p.}
+\textit{Pinter J.D.} Global Optimization: Scientific and Engineering Case Studies. Springer, 2006. 546 p.}
 
-\rlitem{rfa:rulit:lit5}{
-\textit{Стронгин Р.Г., Гергель В.А., Баркалов К.А.} Параллельные методы решения задач глобальной оптимизации. Известия высших учебных заведений. Приборостроение, 2009. Т. 52. № 10. С. 25 – 33.}
+\rlitem{rfa:rulit:litUfa}{
+\textit{Gubaydullin I., Enikeeva L., Barkalov K., Lebedev I.} Parallel Global Search Algorithm for Optimization of the Kinetic Parameters of Chemical Reactions.
+Communications in Computer and Information Science, 2021.
+vol 1510.
+P.
+198–211.}
 
-\rlitem{rfa:rulit:lit6}{
-\textit{Захарова Е.М., Минашина И.К.} Обзор методов многомерной оптимизации. Информационные процессы, 2014. Т 14. №3. С. 256 – 274.}
-
-\rlitem{rfa:rulit:lit7}{
-\textit{Gaviano M., Lera D., Kvasov D.E., Sergeyev Y.D.} Software for generation of classes of test functions with known local and global minima for global optimization.
-ACM Transactions on Mathematical Software.
-– 2003.
-– Vol.
-29.
-– P.
-469-480.}
-
-\rlitem{rfa:rulit:lit8}{
-\textit{Сергеев Я.Д., Квасов Д.Е.} Диагональные методы глобальной оптимизации.  М.: Физматлит, 2008. – 352 c.}
-
-\rlitem{rfa:rulit:lit9}{
-\textit{Gablonsky J.M., Kelley C.T.} A Locally-Biased Form of the DIRECT Algorithm. Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.}
+\rlitem{rfa:rulit:lit16}{
+\textit{Сергеев, Я.Д., Квасов Д.Е.} Диагональные методы глобальной оптимизации. М.: Физматлит, 2008. 352 c.}
 
@@
 487–498.}
+
+\rlitem{rfa:rulit:lit2}{
+\textit{Strongin R.G., Sergeyev Ya.D.} Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, 2000. 704 p.}
+
+\rlitem{rfa:rulit:lit1}{
+\textit{Sergeyev Ya.D., Strongin R.G, Lera D.} Introduction to global optimization exploiting space-filling curves. Springer, 2013. 125 p.}
+
+\rlitem{rfa:rulit:lit3}{
+\textit{Стронгин Р.Г., Гергель В.А., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013. 280 с.}
+
+\rlitem{rfa:rulit:lit5}{
+\textit{Sovrasov V.} Comparison of several stochastic and deterministic derivative-free global optimization algorithms. Lecture Notes in Computer Science, 2019. vol. 11548, P. 70--81.}
+
+%\rlitem{rfa:rulit:lit6}{
+%\textit{Захарова Е.М., Минашина И.К.} Обзор методов многомерной оптимизации. Информационные процессы, 2014. Т 14. №3. С. 256 – 274.}
+
+%\rlitem{rfa:rulit:lit9}{
+%\textit{Gablonsky J.M., Kelley C.T.} A Locally-Biased Form of the DIRECT Algorithm. Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.}
 
@@
 
-\rlitem{rfa:rulit:lit15}{
-\textit{Sysoyev, A., Barkalov, K., Sovrasov, V., Lebedev, I., Gergel, V.} AGS NLP solver. \url{ https://github.com/sovrasov/ags_nlp_solver}}
-
-\rlitem{rfa:rulit:lit16}{
-\textit{Сергеев, Я.Д., Квасов Д.Е.} Диагональные методы глобальной оптимизации. М.: Физматлит, 2008. 352 c.}
+%\rlitem{rfa:rulit:lit15}{ %Эту ссылку можно убрать
+%\textit{Sysoyev, A., Barkalov, K., Sovrasov, V., Lebedev, I., Gergel, V.} AGS NLP solver. \url{ https://github.com/sovrasov/ags_nlp_solver}}
 
@@
 
-\elitem{rfa:rulit:lit1}{
+\elitem{rfa:enlit:lit1}{
 \textit{Sergeyev Ya.D., Strongin R.G, Lera D.} Introduction to global optimization exploiting space-filling curves. Springer, 2013. 125 p.}
 
-\elitem{rfa:rulit:lit2}{
+\elitem{rfa:enlit:lit2}{
 \textit{Strongin R.G., Sergeyev Ya.D.} Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, 2000. 704 p.}
 
-\elitem{rfa:rulit:lit3}{
+\elitem{rfa:enlit:lit3}{
 \textit{Strongin R.G., Gergel V.A., Grishagin V.A., Barkalov K.A.} %Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013. 280 с.
@@
 
-\elitem{rfa:rulit:lit4}{
+\elitem{rfa:enlit:lit4}{
 \textit{Pinter (Ed.), J.D.} Global Optimization: Scientific and Engineering Case Studies. Springer, 2006. 546 p.}
 
-\elitem{rfa:rulit:lit5}{
+\elitem{rfa:enlit:lit5}{
 \textit{Strongin R.G., Gergel V.A., Barkalov K.A.} %Параллельные методы решения задач глобальной оптимизации. Известия высших учебных заведений. Приборостроение, 2009. Т. 52. № 10. С. 25 – 33.
@@
 
-\elitem{rfa:rulit:lit6}{
+\elitem{rfa:enlit:lit6}{
 \textit{Zakharova E.M., Minashina I.K.} %Обзор методов многомерной оптимизации. Информационные процессы, 2014. Т 14. №3. С. 256 – 274.
@@
 
-\elitem{rfa:rulit:lit7}{
+\elitem{rfa:enlit:lit7}{
 \textit{Gaviano M., Lera D., Kvasov D.E., Sergeyev Y.D.} Software for generation of classes of test functions with known local and global minima for global optimization.
@@
 
-\elitem{rfa:rulit:lit8}{
-\textit{Sergeev Ya.D., Kvasov D.E.} %Диагональные методы глобальной оптимизации.  М.: Физматлит, 2008. – 352 c.
-}
+%\elitem{rfa:enlit:lit8}{ Ссылка дублирована с lit16
+%\textit{Sergeev Ya.D., Kvasov D.E.} %Диагональные методы глобальной оптимизации.  М.: Физматлит, 2008. – 352 c.
+%}
 
-\elitem{rfa:rulit:lit9}{
+\elitem{rfa:enlit:lit9}{
 \textit{Gablonsky J.M., Kelley C.T.} A Locally-Biased Form of the DIRECT Algorithm. Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.}
 
-\elitem{rfa:rulit:lit10}{
+\elitem{rfa:enlit:lit10}{
 \textit{Paulavicius R., Zilinskas J, Grothey A.} Parallel branch and bound for global optimization with combination of Lipschitz bounds.
@@
 
-\elitem{rfa:rulit:lit11}{
+\elitem{rfa:enlit:lit11}{
 \textit{Boreskov A.A., Kharlamov A.A., Markovsky N.D., Mikushin D.N., Mortikov E.V., Myltsev A.A., Sakharnykh N.A., Frolov V.A.} %Параллельные вычисления на GPU.
@@
 
-\elitem{rfa:rulit:lit12}{
+\elitem{rfa:enlit:lit12}{
 \textit{Tanenbaum A., Austin T.} Structured computer organization }
 
-\elitem{rfa:rulit:lit13}{
+\elitem{rfa:enlit:lit13}{
 \textit{Komolov D.A., Myalk R.A., Zobenko A.A., Filippov A.S.} %Системы автоматизированного проектирования фирмы Altera MAX+Plus II и Quartus II. Издательство РадиоСофт, 2002. 355 c.
@@
 
-\elitem{rfa:rulit:lit14}{
+\elitem{rfa:enlit:lit14}{
 \textit{Gergel V.P., Barkalov K.A., Sysoyev A.V.} A novel supercomputer software system for solving time-consuming global optimization problems.
@@
 
-\elitem{rfa:rulit:lit15}{
+\elitem{rfa:enlit:lit15}{
 \textit{Sysoyev, A., Barkalov, K., Sovrasov, V., Lebedev, I., Gergel, V.} AGS NLP solver. \url{ https://github.com/sovrasov/ags_nlp_solver}}
 
-\elitem{rfa:rulit:lit16}{
+\elitem{rfa:enlit:lit16}{
 \textit{Sergeev, Ya.D., Kvasov D.E.} %Диагональные методы глобальной оптимизации. М.: Физматлит, 2008. 352 c.

### 2022-10-13 2022/NumMethOneAPI/NMP_authorGuidelinesRu_2022.tex
Частично исправлен список литературы

@@
 Задачи поиска минимума многоэкстремальных функций часто возникают в различных областях науки, таких как химия, физика и т.д.
-В частности, такие задачи возникают при разработке новых лекарственных препаратов [\ref{rfa:rulit:litSulimov}], при поиске конфигураций новых химических соединений [\ref{rfa:rulit:litPosypkin1},\ref{rfa:rulit:litPosypkin2}].
+В частности, такие задачи возникают при разработке новых лекарственных препаратов [\ref{rfa:rulit:lit1}], при поиске конфигураций новых химических соединений [\ref{rfa:rulit:lit2},\ref{rfa:rulit:lit3}].
 Задачи глобальной оптимизации возникают и во многих других приложениях [\ref{rfa:rulit:lit4}].
 Традиционной сферой применения методов глобальной оптимизации стала идентификация параметров математических моделей по данным экспериментов.
-В задачах такого вида требуется провести поиск значений неизвестных параметров модели, при которых результаты расчетов близки к результатам, полученным экспериментально [\ref{rfa:rulit:litUfa}].
+В задачах такого вида требуется провести поиск значений неизвестных параметров модели, при которых результаты расчетов близки к результатам, полученным экспериментально [\ref{rfa:rulit:lit5}].
 
@@
 
-Многие известные алгоритмы липшицевой глобальной оптимизации основаны на идее редукции размерности и адаптации одномерных алгоритмов для решения многомерных задач [\ref{rfa:rulit:lit16},\ref{rfa:rulit:lit10}].
-В данной работе мы будем использовать подход из [\ref{rfa:rulit:lit2}, \ref{rfa:rulit:lit1}], основанный на идее редукции размерности с помощью кривой Пеано  
+Многие известные алгоритмы липшицевой глобальной оптимизации основаны на идее редукции размерности и адаптации одномерных алгоритмов для решения многомерных задач [\ref{rfa:rulit:lit6},\ref{rfa:rulit:lit7}].
+В данной работе мы будем использовать подход из [\ref{rfa:rulit:lit8}, \ref{rfa:rulit:lit9}], основанный на идее редукции размерности с помощью кривой Пеано  
 $y(x)$, которая непрерывно и однозначно отображает отрезок вещественной оси $[0,1]$ на $N$-мерный гиперкуб:
@@
 Пользуясь этим свойством, можно трактовать исходную задачу (\ref{task}) как задачу минимизации одномерной функции $f\left(x\right)=\varphi\left(y\left(x\right)\right)$, $x\in[0,1]$, удовлетворяющей условию Гельдера.
-Для решения указанной задачи минимизации может быть применен \textit{алгоритм глобального поиска} [\ref{rfa:rulit:lit3}], который является эффективным алгоритмом, т.к.
-при решении сложных многоэкстремальных задач опережает (по числу итераций, требующихся для корректного решения задачи с фиксированной точностью) другие методы аналогичного назначения [\ref{rfa:rulit:lit5}].
+Для решения указанной задачи минимизации может быть применен \textit{алгоритм глобального поиска} [\ref{rfa:rulit:lit10}], который является эффективным алгоритмом, т.к.
+при решении сложных многоэкстремальных задач опережает (по числу итераций, требующихся для корректного решения задачи с фиксированной точностью) другие методы аналогичного назначения [\ref{rfa:rulit:lit11}].
 Алгоритм глобального поиска является весьма гибким и отлично подходит для реализации на параллельных вычислительных системах.
@@
 
-Опишем (в соответствии с [\ref{rfa:rulit:lit2}]) теоретические свойства рассматриваемого параллельного алгоритма.
+Опишем (в соответствии с [\ref{rfa:rulit:lit8}]) теоретические свойства рассматриваемого параллельного алгоритма.
 В решаемых задачах время выполнения одного испытания существенно превышает время обработки его результата.
@@
 с нулевой избыточностью) распараллеливанию (доказательство утверждения см.
-в [\ref{rfa:rulit:lit2}]).
+в [\ref{rfa:rulit:lit8}]).
 
@@
 Вместе с тем быстрый прогресс параллельных вычислительных средств предоставляет всё больше различных новых возможностей для решения сложных задач.
-Однако на фоне этого возникает вопрос эффективного распараллеливания, а многообразие различных типов устройств и средств разработки ставят пользователя перед проблемой выбора одной из них [\ref{rfa:rulit:lit10}].
+Однако на фоне этого возникает вопрос эффективного распараллеливания, а многообразие различных типов устройств и средств разработки ставят пользователя перед проблемой выбора одной из них [\ref{rfa:rulit:lit7}].
 
@@
 Несмотря на то, что большинство процессоров сейчас многоядерные, достаточно много ресурсов тратится на обеспечение их однопоточной производительности.
-В GPU большая часть ресурсов задействована под вычисления [\ref{rfa:rulit:lit11}], а не под кэш, как в случае с CPU, поэтому в GPU применима стратегия SIMD[\ref{rfa:rulit:lit12}].
+В GPU большая часть ресурсов задействована под вычисления [\ref{rfa:rulit:lit12}], а не под кэш, как в случае с CPU, поэтому в GPU применима стратегия SIMD[\ref{rfa:rulit:lit13}].
 Третий класс -- это матричные процессоры, рассчитанные на быструю работу с матрицами.
 Прежде всего это ускорители из области искусственного интеллекта, нейронные процессоры, например, процессоры машинного зрения, тензорные и другие.
-И последний класс -- программируемые логические интегральные схемы (ПЛИС), FPGA [\ref{rfa:rulit:lit13}].
+И последний класс -- программируемые логические интегральные схемы (ПЛИС), FPGA [\ref{rfa:rulit:lit14}].
 Основу структуры составляет матрица логических элементов, функции этих элементов и связи между ними могут модифицироваться – программироваться, в процессе использования.
@@
 Для получения исполняемого программного кода использовался компилятор Intel oneAPI DPC++ 2021.1.
-Вычислительные эксперименты выполнялись с использованием  программной системы Globalizer [\ref{rfa:rulit:lit14}].
+Вычислительные эксперименты выполнялись с использованием  программной системы Globalizer [\ref{rfa:rulit:lit15}].
 
@@
 Далее были проведены эксперименты на серии задач, полученных интегрированием функций из генератора GKLS.
-Данный генератор описан в [\ref{rfa:rulit:lit16}].
+Данный генератор описан в [\ref{rfa:rulit:lit6}].
 С его помощью можно порождать задачи многоэкстремальной оптимизации и варьировать их свойства: размерность, количество локальных минимумов, размеры их областей притяжения, координаты точки глобального минимума, значение функции в ней и т.д.
@@
 
-\rlitem{rfa:rulit:litSulimov}{
+\rlitem{rfa:rulit:lit1}{
 \textit{Kutov D., Sulimov A., Sulimov V.} Supercomputer docking: investigation of low energy minima of protein-ligand complexes.
@@
 
-\rlitem{rfa:rulit:litPosypkin1}{
+\rlitem{rfa:rulit:lit2}{
 \textit{Абгарян К.К., Посыпкин М.
@@
 
-\rlitem{rfa:rulit:litPosypkin2}{
+\rlitem{rfa:rulit:lit3}{
 \textit{Евтушенко Ю.Г., Лурье С.А., Посыпкин М.А., Соляев Ю.О.} Применение методов оптимизации для поиска равновесных состояний двумерных кристаллов.
@@
 
-\rlitem{rfa:rulit:litUfa}{
+\rlitem{rfa:rulit:lit5}{
 \textit{Gubaydullin I., Enikeeva L., Barkalov K., Lebedev I.} Parallel Global Search Algorithm for Optimization of the Kinetic Parameters of Chemical Reactions.
@@
 
-\rlitem{rfa:rulit:lit16}{
+\rlitem{rfa:rulit:lit6}{
 \textit{Сергеев, Я.Д., Квасов Д.Е.} Диагональные методы глобальной оптимизации. М.: Физматлит, 2008. 352 c.}
 
-\rlitem{rfa:rulit:lit10}{
+\rlitem{rfa:rulit:lit7}{
 \textit{Paulavicius R., Zilinskas J, Grothey A.} Parallel branch and bound for global optimization with combination of Lipschitz bounds.
@@
 
-\rlitem{rfa:rulit:lit2}{
+\rlitem{rfa:rulit:lit8}{
 \textit{Strongin R.G., Sergeyev Ya.D.} Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, 2000. 704 p.}
 
-\rlitem{rfa:rulit:lit1}{
+\rlitem{rfa:rulit:lit9}{
 \textit{Sergeyev Ya.D., Strongin R.G, Lera D.} Introduction to global optimization exploiting space-filling curves. Springer, 2013. 125 p.}
 
-\rlitem{rfa:rulit:lit3}{
-\textit{Стронгин Р.Г., Гергель В.А., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013. 280 с.}
+\rlitem{rfa:rulit:lit10}{
+\textit{Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013. 280 с.}
 
-\rlitem{rfa:rulit:lit5}{
+\rlitem{rfa:rulit:lit11}{
 \textit{Sovrasov V.} Comparison of several stochastic and deterministic derivative-free global optimization algorithms. Lecture Notes in Computer Science, 2019. vol. 11548, P. 70--81.}
 
-%\rlitem{rfa:rulit:lit6}{
-%\textit{Захарова Е.М., Минашина И.К.} Обзор методов многомерной оптимизации. Информационные процессы, 2014. Т 14. №3. С. 256 – 274.}
-
-%\rlitem{rfa:rulit:lit9}{
-%\textit{Gablonsky J.M., Kelley C.T.} A Locally-Biased Form of the DIRECT Algorithm. Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.}
-
-\rlitem{rfa:rulit:lit11}{
+\rlitem{rfa:rulit:lit12}{
 \textit{Боресков А.А., Харламов А.А., Марковский Н.Д., Микушин Д.Н., Мортиков Е.В., Мыльцев А.А., Сахарных Н.А., Фролов В.А.} Параллельные вычисления на GPU.
@@
 
-\rlitem{rfa:rulit:lit12}{
+\rlitem{rfa:rulit:lit13}{
 \textit{Таненбаум Э., Остин Т.} Архитектура компьютера / перевод с английского Е. Матвеева. — 6-е изд. — Санкт-Петербург [и др.]: Питер, 2014. 816 с.}
 
-\rlitem{rfa:rulit:lit13}{
+\rlitem{rfa:rulit:lit14}{
 \textit{Комолов Д.А., Мяльк Р.А., Зобенко А.А., Филиппов А.С.} Системы автоматизированного проектирования фирмы Altera MAX+Plus II и Quartus II. Издательство РадиоСофт, 2002. 355 c.}
 
-\rlitem{rfa:rulit:lit14}{
+\rlitem{rfa:rulit:lit15}{
 \textit{Gergel V.P., Barkalov K.A., Sysoyev A.V.} A novel supercomputer software system for solving time-consuming global optimization problems.
@@
 
-%\rlitem{rfa:rulit:lit15}{ %Эту ссылку можно убрать
+%\rlitem{rfa:rulit:lit6__}{
+%\textit{Захарова Е.М., Минашина И.К.} Обзор методов многомерной оптимизации. Информационные процессы, 2014. Т 14. №3. С. 256 – 274.}
+
+%\rlitem{rfa:rulit:lit9__}{
+%\textit{Gablonsky J.M., Kelley C.T.} A Locally-Biased Form of the DIRECT Algorithm. Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.}
+
+%\rlitem{rfa:rulit:lit95}{ %Эту ссылку можно убрать
 %\textit{Sysoyev, A., Barkalov, K., Sovrasov, V., Lebedev, I., Gergel, V.} AGS NLP solver. \url{ https://github.com/sovrasov/ags_nlp_solver}}
 
-%\rlitem{rfa:rulit:lit17}{
+%\rlitem{rfa:rulit:lit97}{
 %\textit{Gaviano, M., Lera D., Kvasov D.E., Sergeyev Y.D.} Software for generation of classes of test functions with known local and global minima for global optimization/ M.
@@
 \elitem{rfa:enlit:lit1}{
-\textit{Sergeyev Ya.D., Strongin R.G, Lera D.} Introduction to global optimization exploiting space-filling curves. Springer, 2013. 125 p.}
+\textit{D. Kutov, A. Sulimov, V. Sulimov} Supercomputer docking: investigation of low energy minima of protein-ligand complexes. Supercomputing Frontiers and Innovations 5(3), 134--137 (2018).}
 
 \elitem{rfa:enlit:lit2}{
-\textit{Strongin R.G., Sergeyev Ya.D.} Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, 2000. 704 p.}
+\textit{K.K.
+Abgaryan, M.A.
+Posypkin } Application of optimization methods for solving problems of parametric identification of interatomic interaction potentials.
+Comput.
+Math.
+Math.
+Phys.
+54(12), 1994--2001 (2014).
+[in Russian] }
 
 \elitem{rfa:enlit:lit3}{
-\textit{Strongin R.G., Gergel V.A., Grishagin V.A., Barkalov K.A.} %Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013. 280 с.
-}
+\textit{Yu.G.
+Yevtushenko, S.A.
+Lurie, M.A.
+Posypkin, Yu.O.
+Solyaev } Application of optimization methods to search for equilibrium states of two-dimensional crystals.
+Comput.
+Math.
+Math.
+Phys.
+56(12), 2032--2041 (2016).
+[in Russian]}
 
 \elitem{rfa:enlit:lit4}{
-\textit{Pinter (Ed.), J.D.} Global Optimization: Scientific and Engineering Case Studies. Springer, 2006. 546 p.}
+\textit{J.D. Pinter} Global Optimization: Scientific and Engineering Case Studies. Springer, 2006. 546 p.}
 
 \elitem{rfa:enlit:lit5}{
-\textit{Strongin R.G., Gergel V.A., Barkalov K.A.} %Параллельные методы решения задач глобальной оптимизации. Известия высших учебных заведений. Приборостроение, 2009. Т. 52. № 10. С. 25 – 33.
-}
+\textit{I.
+Gubaydullin, L.
+Enikeeva, K.
+Barkalov, I.
+Lebedev} Parallel Global Search Algorithm for Optimization of the Kinetic Parameters of Chemical Reactions.
+Communications in Computer and Information Science, 2021.
+vol 1510.
+P.
+198–211.}
 
 \elitem{rfa:enlit:lit6}{
-\textit{Zakharova E.M., Minashina I.K.} %Обзор методов многомерной оптимизации. Информационные процессы, 2014. Т 14. №3. С. 256 – 274.
-}
+\textit{Ya.D. Sergeev, D.E. Kvasov } Diagonal Global Optimization Methods  (Fizmatlit, Moscow, 2008) [in Russian]}
 
 \elitem{rfa:enlit:lit7}{
-\textit{Gaviano M., Lera D., Kvasov D.E., Sergeyev Y.D.} Software for generation of classes of test functions with known local and global minima for global optimization.
-ACM Transactions on Mathematical Software.
-– 2003.
-– Vol.
-29.
-– P.
-469-480.}
-
-%\elitem{rfa:enlit:lit8}{ Ссылка дублирована с lit16
-%\textit{Sergeev Ya.D., Kvasov D.E.} %Диагональные методы глобальной оптимизации.  М.: Физматлит, 2008. – 352 c.
-%}
-
-\elitem{rfa:enlit:lit9}{
-\textit{Gablonsky J.M., Kelley C.T.} A Locally-Biased Form of the DIRECT Algorithm. Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.}
-
-\elitem{rfa:enlit:lit10}{
-\textit{Paulavicius R., Zilinskas J, Grothey A.} Parallel branch and bound for global optimization with combination of Lipschitz bounds.
+\textit{R.
+Paulavicius, J.
+Zilinskas, A.
+Grothey } Parallel branch and bound for global optimization with combination of Lipschitz bounds.
 Optimization Methods \& Software,  2011.
@@
 
+\elitem{rfa:enlit:lit8}{
+\textit{R.G. Strongin, Ya.D. Sergeyev } Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, 2000. 704 p.}
+
+\elitem{rfa:enlit:lit9}{
+\textit{Ya.D. Sergeyev,  R.G. Strongin, D. Lera } Introduction to global optimization exploiting space-filling curves. Springer, 2013. 125 p.}
+
+\elitem{rfa:enlit:lit10}{
+\textit{R.G. Strongin, V.P. Gergel, V.A. Grishagin, K.A. Barkalov } Parallel Computations in Global Optimization Problems (Izd. Mosk.
+Gos. Univ., Moscow, 2013.) [in Russian]}
+
 \elitem{rfa:enlit:lit11}{
-\textit{Boreskov A.A., Kharlamov A.A., Markovsky N.D., Mikushin D.N., Mortikov E.V., Myltsev A.A., Sakharnykh N.A., Frolov V.A.} %Параллельные вычисления на GPU.
-Архитектура и программная модель CUDA.
-М.:  Издательство Московского университета, 2015.
-336 с.
-}
+\textit{V. Sovrasov } Comparison of several stochastic and deterministic derivative-free global optimization algorithms. Lecture Notes in Computer Science, 2019. vol. 11548, P. 70--81.}
 
 \elitem{rfa:enlit:lit12}{
-\textit{Tanenbaum A., Austin T.} Structured computer organization }
+\textit{A.A.
+Boreskov, A.A.
+Kharlamov, N.D.
+Markovsky, D.N.
+Mikushin, E.V.
+Mortikov, A.A.
+Myltsev, N.A.
+Sakharnykh, V.A.
+Frolov } Parallel computing on the GPU.
+Architecture and programming model of CUDA.
+(Izd.
+Mosk.
+Gos.
+Univ., Moscow, 2015.) [in Russian]}
 
 \elitem{rfa:enlit:lit13}{
-\textit{Komolov D.A., Myalk R.A., Zobenko A.A., Filippov A.S.} %Системы автоматизированного проектирования фирмы Altera MAX+Plus II и Quartus II. Издательство РадиоСофт, 2002. 355 c.
-}
+\textit{A. Tanenbaum, T. Austin} Structured Computer Organization / translation from English by E. Matveeva. - 6th ed. ( Piter, Saint Petersburg, 2014. ) [in Russian]}
 
 \elitem{rfa:enlit:lit14}{
-\textit{Gergel V.P., Barkalov K.A., Sysoyev A.V.} A novel supercomputer software system for solving time-consuming global optimization problems.
+\textit{D.A. Komolov, R.A. Myalk, A.A. Zobenko, A.S. Filippov} Computer-aided design systems from Altera MAX+Plus II and Quartus II. ( RadioSoft Publishing House, Moscow, 2002.) [in Russian]}
+
+\elitem{rfa:enlit:lit15}{
+\textit{ V.P.
+Gergel, K.A.
+Barkalov, A.V.
+Sysoyev} A novel supercomputer software system for solving time-consuming global optimization problems.
 Numerical Algebra, Control and Optimization, 2018.
@@
 47-62.}
-
-\elitem{rfa:enlit:lit15}{
-\textit{Sysoyev, A., Barkalov, K., Sovrasov, V., Lebedev, I., Gergel, V.} AGS NLP solver. \url{ https://github.com/sovrasov/ags_nlp_solver}}
-
-\elitem{rfa:enlit:lit16}{
-\textit{Sergeev, Ya.D., Kvasov D.E.} %Диагональные методы глобальной оптимизации. М.: Физматлит, 2008. 352 c.
-}
 

### 2022-10-15 2022/NumMethOneAPI/NMP_authorGuidelinesRu_2022.tex
Обновлено оформление

@@
 \Position{лаборант}{laboratory assistant}
-\Orcid{???}
-\Email{yanikolt@gmail.com}
+\Orcid{0000-0003-4894-2497}
+\Email{yanina.silenko@itmm.unn.ru}
 
@@
 \label{task}
-\varphi(y^*)=\min\{\varphi(y):y\in D\}, \; D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}.
+\varphi(y^*)=\min\{\varphi(y):y\in D\}, \; D=\{y\in R^N:a_i\leq x_i\leq{b_i}, 1\leq{i}\leq{N}\}.
 \end{equation}
@@
 \label{lip}
-|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty,
+|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty,
 \end{equation}
@@
 \label{cube}
-\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1}, 1\leqslant i\leqslant N\rbrace=\{y(x): 0\leqslant x\leqslant 1\}.
+\lbrace y\in R^N:-2^{-1}\leq y_i\leq 2^{-1}, 1\leq i\leq N\rbrace=\{y(x): 0\leq x\leq 1\}.
 \end{equation}
@@
 \label{holder}
-\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leqslant H{\left|x_1-x_2\right|}^{\frac{1}{N}}, 
+\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H{\left|x_1-x_2\right|}^{\frac{1}{N}}, 
  x_1,x_2\in[0,1],
@@
 \begin{equation}
-H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}.
+H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leq i\leq N\}.
 \end{equation}
@@
 \begin{equation}
-f_k^*=\min_{1\leqslant i \leqslant k}f(x_i), \; x_k^*=arg \min_{1\leqslant i \leqslant k}f(x_i).
+f_k^*=\min_{1\leq i \leq k}f(x_i), \; x_k^*=arg \min_{1\leq i \leq k}f(x_i).
 \end{equation}
@@
             \end{equation}
-        где $H$ — постоянная Гельдера для одномерной целевой функции.
+        где $H$ --- постоянная Гельдера для одномерной целевой функции.
     \end{enumerate}
@@
 Вместе с тем быстрый прогресс параллельных вычислительных средств предоставляет всё больше различных новых возможностей для решения сложных задач.
-Однако на фоне этого возникает вопрос эффективного распараллеливания, а многообразие различных типов устройств и средств разработки ставят пользователя перед проблемой выбора одной из них [\ref{rfa:rulit:lit7}].
+Однако на фоне этого возникает вопрос эффективного распараллеливания, а многообразие различных типов устройств и средств разработки ставят пользователя перед проблемой выбора одной из них.
 
@@
 В GPU большая часть ресурсов задействована под вычисления [\ref{rfa:rulit:lit12}], а не под кэш, как в случае с CPU, поэтому в GPU применима стратегия SIMD[\ref{rfa:rulit:lit13}].
-Третий класс -- это матричные процессоры, рассчитанные на быструю работу с матрицами.
+Третий класс --- это матричные процессоры, рассчитанные на быструю работу с матрицами.
 Прежде всего это ускорители из области искусственного интеллекта, нейронные процессоры, например, процессоры машинного зрения, тензорные и другие.
-И последний класс -- программируемые логические интегральные схемы (ПЛИС), FPGA [\ref{rfa:rulit:lit14}].
-Основу структуры составляет матрица логических элементов, функции этих элементов и связи между ними могут модифицироваться – программироваться, в процессе использования.
+И последний класс --- программируемые логические интегральные схемы (ПЛИС), FPGA [\ref{rfa:rulit:lit14}].
+Основу структуры составляет матрица логических элементов, функции этих элементов и связи между ними могут модифицироваться --- программироваться, в процессе использования.
 Сфера применения ПЛИС достаточно широка, они используются в бытовой электронике, телекоммуникационном оборудовании, разнообразной робототехнике и при прототипировании микросхем.
@@
   \includegraphics[width=0.7\linewidth]{./pic/s1.pdf}
-  \caption{Общая схема организации параллельных вычислений.}
+  \bicaption{Общая схема организации параллельных вычислений.}{General scheme of organization of parallel computing.}
   \label{fig:s1}  
@@
   \includegraphics[width=1.0\linewidth]{./pic/s2.png}
-  \caption{Линии уровня двумерной функции Растригина (слева) и модифицированной четырехмерной функции Растригина (справа).}
+  \bicaption{Линии уровня двумерной функции Растригина (слева) и модифицированной четырехмерной функции Растригина (справа).} {Level lines of the two-dimensional Rastrigin function (left) and the modified four-dimensional Rastrigin function (right).}
   \label{fig:s2}  
@@
     \centering
-            \caption{Число итераций ПАГП при решении интегрированной функции Растригина.}
+            \bicaption{Число итераций ПАГП при решении интегрированной функции Растригина.}{The number of iterations of PAGP when solving the integrated Rastrigin function.}
     \label{table1}
@@
     \hline
-        N & CPU &   & ~ & ~ & ~ & GPU  & ~ & ~ \\ \hline
+        N &\multicolumn{4}{c|}{CPU} & ~ & \multicolumn{3}{c|}{GPU} \\ \hline
         ~ & P=1 & P=2 & P=4 & P=8 &   & P=256 & P=512 & P=1024  \\ \hline
@@
     \centering
-            \caption{Ускорение ПАГП при решении интегрированной функции Растригина}
+            \bicaption{Ускорение ПАГП при решении интегрированной функции Растригина} {Acceleration of GAGP in solving the integrated Rastrigin function}
     \label{table2}
@@
     \hline
-        N & CPU &   & ~ & ~ & GPU  & ~ & ~ \\ \hline
+        N &\multicolumn{3}{c|}{CPU} & ~ & \multicolumn{3}{c|}{GPU} \\ \hline
         ~ & P=2 & P=4 & P=8 & ~ & P=256 & P=512 & P=1024  \\ \hline
@@
   \includegraphics[width=1.0\linewidth]{./pic/s3.png}
-  \caption{Функции GKLS (слева) и модифицированной функции GKLS (справа).}
+  \bicaption{Функции GKLS (слева) и модифицированной функции GKLS (справа).}{GKLS function (left) and modified GKLS function (right).}
   \label{fig:s3}  
@@
     \centering
-    \caption{Ускорение ПАГП при решении серии модифицированных функций GKLS.}
+    \bicaption{Ускорение ПАГП при решении серии модифицированных функций GKLS.}{Acceleration of GAGP when solving a series of modified GKLS functions.}
     \label{table3}
@@
 
-Полученные результаты подтверждают, что использование инструментов Intel oneAPI для реализации параллельного алгоритма глобального поиска позволяет написать одну версию кода, которая будет будет показывать хорошее ускорение как при использовании центрального процессора, так и при использовании графической карты.
+Полученные результаты подтверждают, что использование инструментов Intel oneAPI для реализации параллельного алгоритма глобального поиска позволяет написать одну версию кода, которая будет показывать хорошее ускорение как при использовании центрального процессора, так и при использовании графической карты.
 
@@
 %
+
 \LITERRUS
@@
 \rlitem{rfa:rulit:lit1}{
-\textit{Kutov D., Sulimov A., Sulimov V.} Supercomputer docking: investigation of low energy minima of protein-ligand complexes.
-Supercomputing Frontiers and Innovations, 2018.
-vol.
+\textit{Kutov D., Sulimov A., Sulimov V.} Supercomputer docking: investigation of low energy minima of protein-ligand complexes ~// Supercomputing Frontiers and Innovations, 2018.
 5(3).
-P.
-134--137.}
+134--137.
+\doi{10.14529/jsfi180326}}
 
@@
 \textit{Абгарян К.К., Посыпкин М.
-А.} Применение оптимизационных методов для решения задач параметрической идентификации потенциалов межатомного взаимодействия.
-Ж.
+А.} Применение оптимизационных методов для решения задач параметрической идентификации потенциалов межатомного взаимодействия ~// Ж.
 вычисл.
@@
 физ., 2014.
-Т.
-54, №12.
-С.1994--2001.}
+\textbf{54}, N~12.
+1994--2001.
+\doi{10.7868/S0044466914120023}}
 
 \rlitem{rfa:rulit:lit3}{
-\textit{Евтушенко Ю.Г., Лурье С.А., Посыпкин М.А., Соляев Ю.О.} Применение методов оптимизации для поиска равновесных состояний двумерных кристаллов.
-Ж.
+\textit{Евтушенко Ю.Г., Лурье С.А., Посыпкин М.А., Соляев Ю.О.} Применение методов оптимизации для поиска равновесных состояний двумерных кристаллов ~// Ж.
 вычисл.
@@
 физ., 2016.
-T.
-56, №12.
-С.
-2032--2041.}
+\textbf{56}, N~12.
+2032--2041.
+\doi{10.7868/S0044466916120097} }
 
 \rlitem{rfa:rulit:lit4}{
-\textit{Pinter J.D.} Global Optimization: Scientific and Engineering Case Studies. Springer, 2006. 546 p.}
+\textit{Pinter J.D.} Global Optimization: Scientific and Engineering Case Studies. Springer, 2006.}
 
 \rlitem{rfa:rulit:lit5}{
-\textit{Gubaydullin I., Enikeeva L., Barkalov K., Lebedev I.} Parallel Global Search Algorithm for Optimization of the Kinetic Parameters of Chemical Reactions.
-Communications in Computer and Information Science, 2021.
-vol 1510.
-P.
-198–211.}
+\textit{Gubaydullin I., Enikeeva L., Barkalov K., Lebedev I.} Parallel Global Search Algorithm for Optimization of the Kinetic Parameters of Chemical Reactions ~// Communications in Computer and Information Science, 2021.
+{\bf 1510}, 198--211.
+\doi{10.1007/978-3-030-92864-3_16}}
 
 \rlitem{rfa:rulit:lit6}{
-\textit{Сергеев, Я.Д., Квасов Д.Е.} Диагональные методы глобальной оптимизации. М.: Физматлит, 2008. 352 c.}
-
-\rlitem{rfa:rulit:lit7}{
-\textit{Paulavicius R., Zilinskas J, Grothey A.} Parallel branch and bound for global optimization with combination of Lipschitz bounds.
-Optimization Methods \& Software,  2011.
-– Vol.
-26, No.
-3.
-P.
-487–498.}
+\textit{Сергеев, Я.Д., Квасов Д.Е.} Диагональные методы глобальной оптимизации. М.: Физматлит, 2008.}
 
 \rlitem{rfa:rulit:lit8}{
-\textit{Strongin R.G., Sergeyev Ya.D.} Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, 2000. 704 p.}
+\textit{Strongin R.G., Sergeyev Ya.D.} Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, 2000.}
 
 \rlitem{rfa:rulit:lit9}{
-\textit{Sergeyev Ya.D., Strongin R.G, Lera D.} Introduction to global optimization exploiting space-filling curves. Springer, 2013. 125 p.}
+\textit{Sergeyev Ya.D., Strongin R.G, Lera D.} Introduction to global optimization exploiting space-filling curves. Springer, 2013.}
 
 \rlitem{rfa:rulit:lit10}{
-\textit{Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013. 280 с.}
+\textit{Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А.} Параллельные вычисления в задачах глобальной оптимизации. М.: Издательство Московского университета, 2013.}
 
 \rlitem{rfa:rulit:lit11}{
-\textit{Sovrasov V.} Comparison of several stochastic and deterministic derivative-free global optimization algorithms. Lecture Notes in Computer Science, 2019. vol. 11548, P. 70--81.}
+\textit{Sovrasov V.} Comparison of several stochastic and deterministic derivative-free global optimization algorithms ~// Lecture Notes in Computer Science, 2019.
+\textbf{11548}.
+70--81.
+\doi{10.1007/978-3-030-22629-9_6} }
 
@@
 Архитектура и программная модель CUDA.
-М.:  Издательство Московского университета, 2015.
-336 с.}
+М.:  Издательство Московского университета, 2015.}
 
 \rlitem{rfa:rulit:lit13}{
-\textit{Таненбаум Э., Остин Т.} Архитектура компьютера / перевод с английского Е. Матвеева. — 6-е изд. — Санкт-Петербург [и др.]: Питер, 2014. 816 с.}
+\textit{Таненбаум Э., Остин Т.} Архитектура компьютера / перевод с английского Е. Матвеева. -- 6-е изд. Санкт-Петербург [и др.]: Питер, 2014.}
 
 \rlitem{rfa:rulit:lit14}{
-\textit{Комолов Д.А., Мяльк Р.А., Зобенко А.А., Филиппов А.С.} Системы автоматизированного проектирования фирмы Altera MAX+Plus II и Quartus II. Издательство РадиоСофт, 2002. 355 c.}
+\textit{Комолов Д.А., Мяльк Р.А., Зобенко А.А., Филиппов А.С.} Системы автоматизированного проектирования фирмы Altera MAX+Plus II и Quartus II. Издательство РадиоСофт, 2002.}
 
 \rlitem{rfa:rulit:lit15}{
-\textit{Gergel V.P., Barkalov K.A., Sysoyev A.V.} A novel supercomputer software system for solving time-consuming global optimization problems.
-Numerical Algebra, Control and Optimization, 2018.
-Vol.
-8, No.
-1.
-P.
-47-62.}
+\textit{Gergel V.P., Barkalov K.A., Sysoyev A.V.} A novel supercomputer software system for solving time-consuming global optimization problems ~// Numerical Algebra, Control and Optimization, 2018.
+\textbf{8}, N~1.
+47--62.
+\doi{10.3934/naco.2018003}}
+
+%\rlitem{rfa:rulit:lit7}{
+%\textit{Paulavicius R., Zilinskas J, Grothey A.} Parallel branch and bound for global optimization with combination of Lipschitz bounds ~// Optimization Methods \& Software,  2011.
+\textbf{26}, N~3.
+487--498.
+\doi{10.1080/10556788.2010.551537} }
 
 %\rlitem{rfa:rulit:lit6__}{
-%\textit{Захарова Е.М., Минашина И.К.} Обзор методов многомерной оптимизации. Информационные процессы, 2014. Т 14. №3. С. 256 – 274.}
+%\textit{Захарова Е.М., Минашина И.К.} Обзор методов многомерной оптимизации. Информационные процессы, 2014. Т 14. №3. С. 256 -- 274.}
 
 %\rlitem{rfa:rulit:lit9__}{
-%\textit{Gablonsky J.M., Kelley C.T.} A Locally-Biased Form of the DIRECT Algorithm. Journal of Global Optimization, – 2001, Vol. 21, No. 1, – P. 27–37.}
+%\textit{Gablonsky J.M., Kelley C.T.} A Locally-Biased Form of the DIRECT Algorithm. Journal of Global Optimization, -- 2001, Vol. 21, No. 1, – P. 27--37.}
 
@@
 \elitem{rfa:enlit:lit1}{
-\textit{D. Kutov, A. Sulimov, V. Sulimov} Supercomputer docking: investigation of low energy minima of protein-ligand complexes. Supercomputing Frontiers and Innovations 5(3), 134--137 (2018).}
+\textit{D.
+Kutov, A.
+Sulimov, V.
+Sulimov} ``Supercomputer docking: investigation of low energy minima of protein-ligand complexes,'' Supercomputing Frontiers and Innovations.
+5(3), 134--137 (2018).
+\doi{10.14529/jsfi180326}}
 
@@
 Abgaryan, M.A.
-Posypkin } Application of optimization methods for solving problems of parametric identification of interatomic interaction potentials.
-Comput.
+Posypkin } ``Application of optimization methods for solving problems of parametric identification of interatomic interaction potentials,'' Comput.
 Math.
@@
 54(12), 1994--2001 (2014).
-[in Russian] }
+[in Russian] \doi{10.7868/S0044466914120023}}
 
@@
 Posypkin, Yu.O.
-Solyaev } Application of optimization methods to search for equilibrium states of two-dimensional crystals.
-Comput.
+Solyaev } ``Application of optimization methods to search for equilibrium states of two-dimensional crystals,'' Comput.
 Math.
@@
 56(12), 2032--2041 (2016).
-[in Russian]}
+[in Russian] \doi{10.7868/S0044466916120097}}
 
 \elitem{rfa:enlit:lit4}{
-\textit{J.D. Pinter} Global Optimization: Scientific and Engineering Case Studies. Springer, 2006. 546 p.}
+\textit{J.D. Pinter} Global Optimization: Scientific and Engineering Case Studies. Springer, 2006.}
 
@@
 Barkalov, I.
-Lebedev} Parallel Global Search Algorithm for Optimization of the Kinetic Parameters of Chemical Reactions.
-Communications in Computer and Information Science, 2021.
-vol 1510.
-P.
-198–211.}
+Lebedev} ``Parallel Global Search Algorithm for Optimization of the Kinetic Parameters of Chemical Reactions,'' Communications in Computer and Information Science.
+1510, 198--211 (2021).
+\doi{10.1007/978-3-030-92864-3_16}}
 
@@
 
-\elitem{rfa:enlit:lit7}{
-\textit{R.
-Paulavicius, J.
-Zilinskas, A.
-Grothey } Parallel branch and bound for global optimization with combination of Lipschitz bounds.
-Optimization Methods \& Software,  2011.
-– Vol.
-26, No.
-3.
-P.
-487–498.}
-
 \elitem{rfa:enlit:lit8}{
-\textit{R.G. Strongin, Ya.D. Sergeyev } Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, 2000. 704 p.}
+\textit{R.G. Strongin, Ya.D. Sergeyev } Global Optimization with Non-convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, 2000.}
 
 \elitem{rfa:enlit:lit9}{
-\textit{Ya.D. Sergeyev,  R.G. Strongin, D. Lera } Introduction to global optimization exploiting space-filling curves. Springer, 2013. 125 p.}
+\textit{Ya.D. Sergeyev,  R.G. Strongin, D. Lera } Introduction to global optimization exploiting space-filling curves. Springer, 2013.}
 
@@
 \elitem{rfa:enlit:lit11}{
-\textit{V. Sovrasov } Comparison of several stochastic and deterministic derivative-free global optimization algorithms. Lecture Notes in Computer Science, 2019. vol. 11548, P. 70--81.}
+\textit{V.
+Sovrasov } ``Comparison of several stochastic and deterministic derivative-free global optimization algorithms,'' Lecture Notes in Computer Science.
+11548, 70--81 (2019).
+\doi{10.1007/978-3-030-22629-9_6}}
 
@@
 \elitem{rfa:enlit:lit13}{
-\textit{A. Tanenbaum, T. Austin} Structured Computer Organization / translation from English by E. Matveeva. - 6th ed. ( Piter, Saint Petersburg, 2014. ) [in Russian]}
+\textit{A. Tanenbaum, T. Austin} Structured Computer Organization / translation from English by E. Matveeva. -- 6th ed. ( Piter, Saint Petersburg, 2014.) [in Russian]}
 
@@
 Barkalov, A.V.
-Sysoyev} A novel supercomputer software system for solving time-consuming global optimization problems.
-Numerical Algebra, Control and Optimization, 2018.
-Vol.
-8, No.
-1.
-P.
-47-62.}
+Sysoyev} ``A novel supercomputer software system for solving time-consuming global optimization problems,'' Numerical Algebra, Control and Optimization.
+8(1), 47--62 (2018).
+\doi{10.3934/naco.2018003}}
+
+%\elitem{rfa:enlit:lit7}{
+%\textit{R.
+Paulavicius, J.
+Zilinskas, A.
+Grothey } ``Parallel branch and bound for global optimization with combination of Lipschitz bounds,'' Optimization Methods \& Software.
+26(3), 487--498 (2011).
+\doi{10.1080/10556788.2010.551537} }
 

### 2022-10-15 2022/NumMethOneAPI/NMP_authorGuidelinesRu_2022.tex
Поправлено англоязычное название
fix one reference

@@
 \Article{О реализации параллельного алгоритма глобальной оптимизации с использованием набора инструментов Intel oneAPI
-}%Обновить перевод названия
+}
     {On implementation of the parallel global optimization algorithm with the Intel oneAPI toolkit}
@@
 Приведены результаты вычислительных экспериментов, полученные при решении серии сложных задач многоэкстремальной оптимизации.}
-%обновить перевод - сейчас он для старой аннотации
 {The paper considers the parallel global optimization algorithm ans discusses its implementation with the Intel oneAPI toolkit.
@@
 \Acknowledgements{Работа выполнена при поддержке программы Центра компетенций OneAPI в ННГУ, Министерства науки и высшего образования РФ (проект № 0729-2020-0055) и научно-образовательного математического центра «Математика технологий будущего» (проект № 075-02-2021-1394).}
-    {Работа выполнена при поддержке программы Центра компетенций OneAPI в ННГУ, Министерства науки и высшего образования РФ (проект № 0729-2020-0055) и научно-образовательного математического центра «Математика технологий будущего» (проект № 075-02-2021-1394).}
+{The work was supported by the oneAPI Center of Excellence, by the Ministry of Science and Higher Education of the Russian Federation (project no.
+0729-2020-0055), and by the Research and Education Mathematical Center (project no.
+075-02-2020-1483/1).}
 
@@
 \YEAR{2022}
-\Received{16 октября 2022 г.}{October 16, 2022}
+\Received{17 октября 2022 г.}{October 17, 2022}
 \Accepted{?? ??????? 2022 г.}{??????? ??, 2022}
@@
 
-Задачи поиска минимума многоэкстремальных функций часто возникают в различных областях науки, таких как химия, физика и т.д.
+Задачи поиска минимума многоэкстремальных функций часто возникают в различных областях науки --- химии, физике, биологии, и т.д.
 В частности, такие задачи возникают при разработке новых лекарственных препаратов [\ref{rfa:rulit:lit1}], при поиске конфигураций новых химических соединений [\ref{rfa:rulit:lit2},\ref{rfa:rulit:lit3}].
@@
 
-Будем предполагать, что информация о структурных свойствах целевой функции $\varphi(y)$ неизвестна, а сама функция задается как <<черный ящик>>, т.е.
+Будем предполагать, что информация о структуре целевой функции $\varphi(y)$ неизвестна, а сама функция задается как <<черный ящик>>, т.е.
 с помощью некоторого реализованного программно алгоритма вычисления ее значений.
@@
 Многие известные алгоритмы липшицевой глобальной оптимизации основаны на идее редукции размерности и адаптации одномерных алгоритмов для решения многомерных задач [\ref{rfa:rulit:lit6},\ref{rfa:rulit:lit7}].
-В данной работе мы будем использовать подход из [\ref{rfa:rulit:lit8}, \ref{rfa:rulit:lit9}], основанный на идее редукции размерности с помощью кривой Пеано  
+В данной работе мы будем использовать подход из [\ref{rfa:rulit:lit8},\ref{rfa:rulit:lit9}], основанный на идее редукции размерности с помощью кривой Пеано  
 $y(x)$, которая непрерывно и однозначно отображает отрезок вещественной оси $[0,1]$ на $N$-мерный гиперкуб:
@@
     \centering
-            \bicaption{Число итераций ПАГП при решении интегрированной функции Растригина.}{The number of iterations of PAGP when solving the integrated Rastrigin function.}
+            \bicaption{Число итераций ПАГП при минимизации модифицированной функции Растригина.}{The number of iterations of PAGP when minimizing modified Rastrigin function.}
     \label{table1}
@@
     \centering
-            \bicaption{Ускорение ПАГП при решении интегрированной функции Растригина} {Acceleration of GAGP in solving the integrated Rastrigin function}
+            \bicaption{Ускорение ПАГП при минимизации модифицированной функции Растригина} {Acceleration of GAGP when miminimizing modified Rastrigin function}
     \label{table2}
@@
 
-\begin{figure}
+\begin{figure}[!ht]
 \begin{center}
@@
 
+\rlitem{rfa:rulit:lit7}{
+\textit{Paulavicius R., Zilinskas J.} Simplicial Global Optimization. Springer, New York. 2014. \doi{10.1007/978-1-4614-9093-7} }
+
 \rlitem{rfa:rulit:lit8}{
@@
 \doi{10.3934/naco.2018003}}
-
-%\rlitem{rfa:rulit:lit7}{
-%\textit{Paulavicius R., Zilinskas J, Grothey A.} Parallel branch and bound for global optimization with combination of Lipschitz bounds ~// Optimization Methods \& Software,  2011.
-\textbf{26}, N~3.
-487--498.
-\doi{10.1080/10556788.2010.551537} }
 
@@
 Lurie, M.A.
-Posypkin, Yu.O.
-Solyaev } ``Application of optimization methods to search for equilibrium states of two-dimensional crystals,'' Comput.
+Posypkin, Y.O.
+Solyaev } ``Application of optimization methods for finding equilibrium states of two-dimensional crystals,'' Comput.
 Math.
@@
 \textit{Ya.D. Sergeev, D.E. Kvasov } Diagonal Global Optimization Methods  (Fizmatlit, Moscow, 2008) [in Russian]}
+
+\elitem{rfa:enlit:lit7}{
+\textit{Paulavicius R., Zilinskas J.} Simplicial Global Optimization. Springer, New York. 2014. \doi{10.1007/978-1-4614-9093-7} }
 

### 2022-10-15 2022/NumMethOneAPI/NMP_authorGuidelinesRu_2022.tex
fix brackets

@@
 И последний класс --- программируемые логические интегральные схемы (ПЛИС), FPGA [\ref{rfa:rulit:lit14}].
-Основу структуры составляет матрица логических элементов, функции этих элементов и связи между ними могут модифицироваться --- программироваться, в процессе использования.
+Основу структуры составляет матрица логических элементов, функции этих элементов и связи между ними могут модифицироваться (программироваться) в процессе использования.
 Сфера применения ПЛИС достаточно широка, они используются в бытовой электронике, телекоммуникационном оборудовании, разнообразной робототехнике и при прототипировании микросхем.

### 2022-10-15 2022/NumMethOneAPI/NMP_authorGuidelinesRu_2022.tex
fix oneAPI

@@
 The parallel algorithm presented in the paper uses the scheme of dimensionality reduction employing the Peano curve, which continuously maps an interval of the real axis onto a hypercube.
-The Intel OneApi tools, that allows one to write the same code for both the central processor and the graphics accelerator, were used for implementation of the parallel global optimization algorithm.
+The Intel oneAPI tools, that allows one to write the same code for both the central processor and the graphics accelerator, were used for implementation of the parallel global optimization algorithm.
 The results of numerical experiments obtained by solving a series of time-consuming multiextremal optimization problems are presented.}
@@
 графические ускорители,
-Intel OneAPI.}
+Intel oneAPI.}
         {global optimization,
@@
 graphics accelerators,
-Intel OneApi.}
+Intel oneAPI.}
 
-\Acknowledgements{Работа выполнена при поддержке программы Центра компетенций OneAPI в ННГУ, Министерства науки и высшего образования РФ (проект № 0729-2020-0055) и научно-образовательного математического центра «Математика технологий будущего» (проект № 075-02-2021-1394).}
+\Acknowledgements{Работа выполнена при поддержке программы Центра компетенций oneAPI в ННГУ, Министерства науки и высшего образования РФ (проект № 0729-2020-0055) и научно-образовательного математического центра «Математика технологий будущего» (проект № 075-02-2021-1394).}
 {The work was supported by the oneAPI Center of Excellence, by the Ministry of Science and Higher Education of the Russian Federation (project no.

### 2023-01-12 2023/PAVT/ParallelDecisionTree.tex
Первая версия статьи на PAVT по деревьям решений

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{svproc}
\usepackage{marvosym}
%
% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\usepackage{graphicx}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\usepackage{hyperref}
\usepackage[english, russian]{babel} 
\usepackage{amsmath} 
\def\UrlFont{\rmfamily}
\providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}

\def\orcidID#1{\unskip$^{[#1]}$}
\def\letter{$^{\textrm{(\Letter)}}$}

\begin{document}
\mainmatter              % start of a contribution
%
\title{Combining Global and Local Optimization Algorithms Using Decision Trees}
%
\titlerunning{Global Optimization Using Decision Trees}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{K.~A.~Barkalov\inst{1}\letter\orcidID{0000-0001-5273-2471} \and I.~G.~Lebedev\inst{1}\letter\orcidID{0000-0002-8736-0652} \and D.~I.~Silenko\inst{1}\letter\orcidID{0000-0002-2578-9699}}
%
\authorrunning{K.~A.~Barkalov et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{K.~A.~Barkalov and I.~G.~Lebedev, and D.~I.~Silenko}
%
\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
\email{konstantin.barkalov@itmm.unn.ru, ilya.lebedev@itmm.unn.ru, rotor12587@mail.ru}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
В работе рассматривается решение задач многомерной глобальной оптимизации с применением decision tree для выявления областей притяжения локальных минимумов. Искомая функция задана как черный ящик и может быть выдвинуто лишь общее предположение о том, что она удовлетворяет условию Липшица с неизвестной константой. 
В ходе решения задачи алгоритмом глобального поиска производится анализ позволяющий принять решение о запуске локального метода. Локальные методы играют роль локального уточнения и призваны ускорить сходимость алгоритма. Для подтверждения теории, были приведены вычислительных экспериментов, демонстрирующие ускорение при решении серии тестовых задач. \dots
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$  Decision tree}
\end{abstract}
%

\section{Введение}

При поиске глобального минимума функции можно придерживаться различных алгоритмов: от основанных на идее случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3}, до детерминированных алгоритмов гарантирующих сходимость к глобальному минимуму \cite{fio_bib4, fio_bib5, fio_bib6}. Поскольку в реальных задачах глобальной оптимизации каждое вычисление значения функции представляет собой весьма трудоемкую задачу, необходимо уменьшить количество таких операций. Этого можно добиться целенаправленным выбором вариантов в процессе поиска оптимального решения. На этой идее основывается алгоритм глобального поиска (АГП) \cite{fio_bib7}. В данной работе мы попробуем объединить АГП и метод Хука-Дживса (метод локальной оптимизации) с помощью decision tree, с целью уменьшить количество проводимых итераций, а вместе с этим и количество вычислений целевой функции.

Алгоритмы поиска локального экстремума предназначены для определения лишь одного из локальных экстремумов на множестве допустимых решений, в котором целевая функция принимает максимальное или минимальное значение. При их построении могут использоваться как детерминированный спуск в область экстремума, так и случайный поиск. Среди детерминированных различают методы нулевого порядка и градиентные (1-го и 2-го порядка) \cite{fio_bib8, fio_bib9}. В данной работе использовался метод Хука-Дживса.

Для того, чтобы понять, в какой момент лучше всего запускать локальный метод, мы используем алгоритм на основе decision tree. Мы накапливаем определенное число точек и лишь когда их станет достаточно начинаем обработку. Эти полученные точки мы используем для тренировки decision tree и предсказания по нему. Предсказание позволяет нам получить кусочно-линейную аппроксимацию по этому дереву. А дальше мы рассматриваем уже полученную по дереву аппроксимацию. Проверяем значения соседних точек к той, которую считаем подозрительной на минимум, и оцениваем попали мы в локальную ямку или нет. Причем, decision tree можно обобщить на многомерный случай (подробнее об этом расскажем дальше).

\section{Постановка задачи}

Рассмотрим задачу поиска глобального минимума функции $\varphi(y)$ в гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1\le\ i\le\ n \}$ При этом будем предполагать, что функция удовлетворяет условию Липшица с априори неизвестной константой $L$. Что соответствует ограниченности изменения значений функции при ограниченной вариации аргумента. Это предположение можно интерпретировать (применительно к прикладным задачам) как отражение ограниченности мощностей, порождающих изменения в моделируемой системе.

\begin{equation} \label{sec:problem}   
\varphi(y^*) = min\{\varphi(y):y\in D\}, D = \{y \in R^N : a_i \leq y_i \leq b_i, 1 \leq i \leq N \},
\end{equation}
где $a,b \in R$ --- заданные векторы.

\begin{displaymath}
|\varphi(y_1)-\varphi(y_2)|\leq L\parallel y_1-y_2 \parallel
,y_1,y_2 \in D, 0<L< \infty.
\end{displaymath}

Численное решение задачи (\ref{sec:problem})   сводится к построению оценки $ y_k^\ast\in\ D$ , отвечающей некоторому понятию близости к точке $y^\ast$ (например, ${||y^\ast-y}_k^\ast||\le\ \varepsilon$, где $\varepsilon\geq0$ есть заданная точность) на основе конечного числа $k$ вычислений значений оптимизируемой функции. Относительно класса рассматриваемых задач предполагается, что оптимизируемая функция $\varphi(y)$ может быть задана не аналитически, в виде формулы, а алгоритмически, как результат работы некоторой подпрограммы или библиотеки.

Задачи многоэкстремальной оптимизации имеют существенно более высокую трудоемкость решения по сравнению с другими типами оптимизационных задач, т.к. глобальный оптимум является интегральной характеристикой решаемой задачи и требует исследования всей области поиска. Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (сетки) в области параметров, и выборе наилучшего значения функции на данной сетке. Вычислительные затраты на решение задачи растут экспоненциально с ростом размерности.

В ННГУ им. Н.И. Лобачевского под руководством проф. Р.Г. Стронгина разработан эффективный подход к решению задач глобальной оптимизации. В рамках данного подхода решение многомерных задач сводится к решению серии вложенных задач меньшей размерности. Для эффективного решения быстро вычисляемых задач предлагается полностью перенести решение вложенной задачи на графический ускоритель.

\section{Многомерный алгоритм глобального поиска}

Основная идея данного подхода заключается в том, что минимизируемая функция рассматривается как реализация некоторого случайного процесса. Решающие правила алгоритма конструируются таким образом, что очередная итерация проводится в точке глобального минимума математического ожидания значений функции. Эта точка записывается в список известных значений и итерации повторяются. Так происходит до тех пор, пока не достигнут один из выбранных критериев остановки: расстояние между точками отрезка не становится меньше заданного значения или новые точки не попадают в окрестность истинного глобального минимума. Еще одним вариантом остановки работы алгоритма является достижение установленного максимума возможных итераций. \cite{fio_bib10}

При решении многомерных задач применяется редукция размерности (т.е. сведение многомерной задачи к эквивалентной одномерной) с использованием кривых Пеано. Они позволяют свести многомерную задачу минимизации в области $D$  к одномерной задаче минимизации на отрезке $[0, 1]$

\begin{displaymath}
\varphi(y(x^\ast))\ =\ min\{\varphi(y(x)):\ x\epsilon[0,\ 1]\}
\end{displaymath}

где функция $\varphi(y(x^\ast))$  удовлетворяют уже равномерному условию Гельдера

\begin{displaymath}
\left|\varphi (y \left(x_1\right))- \varphi (y \left(x_2\right)\right )|\le\ H\left|x_1-x_2\right|^\frac{1}{N},\ x_1,\ x_2\epsilon[0,1]		
\end{displaymath}

Поэтому, не ограничивая общности, можно рассматривать минимизацию одномерной функции $f(x)\ =\ \varphi(y(x)), \ x\epsilon [0,1]$, удовлетворяющей условию Гельдера.

Схема алгоритма:

На подготовительном шаге параллельно проводятся $p$ trials в произвольных внутренних точках $x^1, ...,x^p$ отрезка $[0,1]$, что соответствует первой итерации  алгоритма.

Если выполнено $n\geq1$ итераций, которым соответствуют $k=k(n)$ проведенных trial in point $x^i, 1\leq i\leq k$, то trial $x^{k+1},\ldots,x^{k+p}$  следующей $(n+1)$-й итерации будут определять следующим образом.

\begin{enumerate}

\item  Перенумеровать (нижним индексом) trial ранее проведенных  $x^i, 1\leq i\leq k$, а также граничные точки отрезка [0,1] в порядке возрастания координаты:
 \begin{equation}
\label{agp1_sort}
	0=x_0<\ x_1<\ ...\ <x_{k+1}=1.
	\end{equation}
	и сопоставить им значения $z_i=f(x_i)$.

\item  Вычислить текущие нижние оценки $M$ неизвестной константы Гельдера $H$:
 \begin{equation}
\label{agp2_mu}
	\mu=max\left\{\frac{|z_i-z_{i-1}|}{{{(x}_i-x_{i-1})}^{1/N}},\ i=1,\ldots,k\right\},\ M=\ \left\{\begin{matrix}r\mu,\ \mu>0,\\1,\ \mu=0,\\\end{matrix}\right.\
	\end{equation}
где $r>1$ --- заданный параметр алгоритма.

\item  Для каждого интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить величину $R(i)$, называемую \textit{характеристикой} интервала, в соответствии с формулами

\begin{equation}
\label{agp3_R1}
R(1)=2\Delta_1-4\dfrac{z_1}{M}, \; R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
\end{equation}

\begin{equation}
%\label{agp3_Ri}
R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1,
\end{equation}
где \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).

\item   Упорядочить характеристики $R\left(i\right),\ 1\leq i \leq k+1,$ в порядке невозрастания 
\begin{equation}
\label{agp4_R_sort}
	R\left(t_1\right)\geq\ R\left(t_2\right)\geq...\geq\ R\left(t_k\right)\geq\ R(t_{k+1}),\ 
\end{equation}	
и выбрать $p$ интервалов с номерами $t_j,\ 1\le\ j\le\ p$, с наибольшими значениями характеристики.

\item В выбранных интервалах вычислить точки $x^{k+j},\ 1\leq j\leq p,$ в соответствии с формулами
\begin{equation}
\label{agp5_x1}
	x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\ t_j=1,\ t_j=k+1,
\end{equation}	
	\begin{equation}
\label{agp4_xi}	
	x^{k+1}=\frac{x_{t_j}+x_{t_j-1}}{2}-sign\left(z_{t_j}-z_{t_j-1}\right)\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N,\ 1<t_j<k+1.
\end{equation}

\end{enumerate}

Очередные trial проводятся параллельно, в вычисленных по формулам (\ref{agp5_x1}), (\ref{agp4_xi}) точках $x^{k+j},\ 1\leq j\leq p$.

Отметим, что в прикладных оптимизационных задачах процесс проведения trial является, как правило, значительно более трудоемким по сравнению с работой вычислительных правил алгоритма.

Алгоритм останавливает свою работу в случае, если хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}) выполняется условие \(\Delta_{t_j} < \varepsilon\). Данный критерий остановки (наряду с обычным для итерационных методов критерием, ограничивающим число выполненных итераций) используется в задачах оптимизации, в которых точка глобального минимума $x^*$ заранее неизвестна.

При решении тестовых задач, в которых точка глобального минимума $x^*$ является  известной, можно использовать также и критерий остановки по попаданию в окрестность глобального минимума. В этом случае метод прекращает работу, если если хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}) выполняется условие $\left|x_{t_j}-\ x^\ast\right| < \varepsilon.$

В качестве итоговой оценки глобально-оптимального решения рассматриваемой задачи выбираются значения 
\begin{equation}
f_k^*=\min_{1\leq i \leq k}f(x_i), \; x_k^*=arg \min_{1\leq i \leq k}f(x_i).
\end{equation}

Обоснование данного способа организации вычислений см. в \cite{fio_bib20}. Модификации, учитывающие наличие ограничений-неравенств в задаче, а также информация о производной целевой функции, представлены в \cite{fio_bib12, fio_bib9, fio_bib11}.

\section{Метод Хука-Дживса}

Метод Хука–Дживса – это комбинация исследующего поиска по направлениям и поиска по образцу \cite{fio_bib14, fio_bib15}.

Исследующий поиск определяется следующим образом: 
\begin{enumerate}
\item	Определяется значение шага (оно различно для каждого направления координат и может измениться во время поиска). 
\item	Шаг поиска считается успешным, если значение целевой функции в контрольной точке не превышает значение исходной целевой функции. 
\item	В противном случае нужно вернуться к предыдущему пункту и сделать шаг в обратном направлении. 
\item	После перебора всех N координат исследовательский поиск заканчивается. Полученная точка называется базовой точкой.
\end{enumerate}

Что касается поиска по образцу, то он заключается в реализации шага из найденной в ходе исследующего поиска базовой точки вдоль прямой, соединяющей ее с предыдущей базовой точкой. Величина этого шага определяется задаваемым заранее параметром.

\begin{figure}[!h]
	\begin{center}
		\begin{minipage}[h]{0.8\linewidth}
			\includegraphics[width=1\linewidth]{figure/fig1.png}
			\caption{Пример итераций алгоритма Хука-Дживса} %% подпись к рисунку
			\label{fig:fig1}
		\end{minipage}
	\end{center}
\end{figure}

\section{Decision tree}

Decision tree — это метод автоматического анализа больших массивов данных, который применяется в машинном обучении. Оно представляет собой бинарное дерево (дерево, в котором каждый нелистовой узел имеет два дочерних узла). Его можно использовать как для решения задач классификации, так и для регрессионных задач. В первом случае, каждый лист дерева помечается меткой класса, а несколько листьев могут иметь одну и ту же метку. Во втором же, каждому листу дерева присваивается константа, поэтому получаемая функция аппроксимации является кусочно-постоянной.

Процедура прогнозирования по decision tree начинается с корня. От каждого нелистового узла процедура идет влево или вправо на основании значения некоторой переменной, индекс которой хранится в наблюдаемом узле. Возможны следующие переменные:
Упорядоченные переменные, значения которых сравнивается с порогом (который также хранится в узле). Если значение меньше порога, процедура идет влево. В противном случае - вправо.
Категориальные переменные - значение дискретной переменной проверяется на предмет того, принадлежит ли оно определенному подмножеству значений (также хранящемуся в узле) из набора допустимых величин. Если принадлежит, то идем влево. В противном случае - вправо.

Как только алгоритм достигает конечного узла, в качестве ответа процедуры прогнозирования выбирается значение, присвоенное этому узлу.

Здесь мы описали все лишь в общих чертах и не вдавались в подробности. Это связано с тем, что непосредственно сами decision tree мы не реализовывали, а лишь использовали готовые из библиотеки OpenCV (это библиотека алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения с открытым кодом). Подробнее с decision tree можно ознакомиться в \cite{fio_bib16}

\section{Объединение метода Хука-Дживса и АГП в случае многомерных задач}

Пришла пора описать как использовать decision tree в многомерном случае. В процессе работы АГП мы хотим понять, стоит ли использовать текущую точку, в качестве стартовой для локального метода или это будет бессмысленно. Для этого можно осмотреть соседние точки и если среди них нет ни одной такой, значение функции в которой меньше, чем в текущей, то можно предположить, что мы находимся в подобии локальной ямки и можно запустить локальный метод, чтобы быстрее сойтись к минимуму. Использование проекций многомерных точек на одномерный отрезок (то, что позволяют сделать кривые Пеано) нам для этих целей не подойдут сразу по нескольким причинам. Во-первых, вид функции после отображения очень сильно изменится и искать минимумы станет уже неэффективной идеей. Во-вторых, после отображения мы можем потерять информацию о взаимном расположении точек исходной функции (точки, расположенные близко в многомерном пространстве, могут оказаться на разных концах одномерного отрезка и наоборот).

\begin{figure}[ht!]
	\begin{center}
		\begin{minipage}[h]{0.8\linewidth}
			\includegraphics[width=1\linewidth]{figure/fig2.png}
			\caption{Функция из генератора gkls до использования кривых Пеано} %% подпись к рисунку
			\label{fig:fig2}
		\end{minipage}
	\end{center}
\end{figure}

\begin{figure}[ht!]
	\begin{center}
		\begin{minipage}[h]{0.8\linewidth}
			\includegraphics[width=1\linewidth]{figure/fig2.png}
			\caption{Рисунок с двумерным decision tree} %% подпись к рисунку
			\label{fig:fig2_2}
		\end{minipage}
	\end{center}
\end{figure}

Поэтому мы будем работать именно с исходными точками. Но и тут все не так однозначно. Определить соседние точки в многомерном пространстве достаточно трудоемко, но тут нам как раз и помогут decision tree. Итак, перейдем к самому методу.

После проведения определенного числа trial (например, $100\ \ast\ N$), используем все накопившиеся точки для инициализации подходящей структуры данных и тренировки на ее основе decision tree. Чтобы проще было определить соседние точки, мы построим равномерную сетку с определенным шагом по каждой из размерностей. Но вычислять значение функции в каждой из точек полученной сетки было бы слишком трудозатратно, поэтому мы подадим их на вход предсказанию по дереву. Теперь, имея равномерную сетку точек и зная значение функции в каждой из них, найдем ближайшую, с точки зрения евклидова расстояния, к исходной. Эта ближайшая точка – это, по сути, проекция изначальной точки на равномерную сетку и определить ее соседей уже не такая сложная задача (мы можем просто перебрать индексы)

При рассмотрении соседей необходимо учитывать следующее: если хоть один из соседей имеет значение функции меньшее, чем в текущей точке, то смысла запускать локальный метод нет; если какой-то из соседей имеет значение функции такое же, как в текущей, то его также необходимо проверить (найти всех его соседей и проверять их полностью аналогично); и только в том случае, если все значения функций соседей больше (включая соседей равных точек) мы можем запускать локальный метод.

На рисунке  (\ref{fig:fig2}) отображены линии уровней оптимизационной задачи с отображением текущих точек trial, на рисунен  (\ref{fig:fig2_2}) отображено соответствующее decision tree.

\section{Асинхронная схема распараллеливания с использованием MPI }

Идея асинхронной схемы в том, чтобы не дожидаться окончания вычисления всей отправленной группы точек, а отправлять новую точку, как только была закончена работа с первой из старых.

Предположим, что главный процесс вычисляет одну точку следующего trial на каждой итерации и отправляет ее рабочему процессу для выполнения trial. В то же время выполнение trial рабочим процессом – гораздо более затратная в вычислительном отношении операция, чем выбор мастером новой trial, исключающей простаивающие рабочие процессы. В этом случае (в отличие от синхронных параллельных алгоритмов) общее количество trial, выполняемых каждым рабочим процессом, будет зависеть от вычислительных затрат на выполнение конкретного trial и не может быть оценено заранее. При описании параллельного алгоритма предположим, что в нашем распоряжении имеется $p+1$ вычислительных процессов: один главный и $p$ рабочих.

 В начале поиска главный процесс (допустим, это процесс $0$) инициирует параллельное выполнение $p$ trial в $p$ разных точках области поиска.

Две из этих точек являются граничными, а остальные – внутренними, т. е. в точках $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$, где
$x^1=,x^p=,x^i\in\left(0,1\right),i=2,\ldots,p$

Теперь предположим, что выполнено $k\geq0$ trial и рабочие процессы выполняют trial в точках $y\left(x^{k+1}\right),\;y\left(x^{k+2}\right),\ldots,\;y\left(x^{k+p}\right)$.

Каждый рабочий процесс, завершив в какой-то момент вычисление функции, отправляет главному процессу значение. В свою очередь, главный процесс выбирает новую пробную точку $x^{k+p+1}$ для рабочего процесса в соответствии с описанными ниже правилами. Заметим, что в этом случае мы будем иметь набор прообразов пробных точек 
$I_k=\left\{x^{k+1},x^{k+2},\ldots,x^{k+p}\right\}$
на котором trial уже начаты, но еще не завершены.

Таким образом параллельный ассинхронный алгоритм глобальной оптимизации с использованием decision tree для выявления локальных областей состоит из следующих шагов:

\begin{enumerate}
\item  Перенумеровать множество прообразов точек trial

\begin{displaymath}
X_k=\left\{x^1,x^2,\ldots,x^{k+p}\right\},
\end{displaymath}

содержащие все прообразы, на которых либо завершились, либо идут trial в порядке возрастания (по нижнему индексу) так, что

\begin{displaymath}
0=x_1<x_2<\ldots<x_{k+p}=1.
\end{displaymath}

\item  Вычислите значения

\begin{displaymath}
M_1=\max{\left\{\frac{\left|z_i-z_{i-1}\right|}{\left(x_i-x_{i-1}\right)^{1/N}}:x_{i-1}\notin I_k,x_i\notin I_k,2\le i\le k+p\right\}},
\end{displaymath}

\begin{displaymath}
M_2=\max{\left\{\frac{\left|z_{i+1}-z_{i-1}\right|}{\left(x_{i+1}-x_{i-1}\right)^{1/N}}:x_i\in I_k,2\le i<k+p\right\}},
\end{displaymath}

\begin{displaymath}
M=\max{\{}M_1,M_2\},
\end{displaymath}

где $ z_i=\varphi\left(y\left(x_i\right)\right)$, если $x_i\notin I_k,\;1\le i\le k+p$. Значения $z_i$ в точках $x_i\in I_k$ не определены, так как trial в точках $x_i\in I_k$ еще не завершены. Если значение $M$ равно $0$, то установите $M=1$.

\item  Сопоставьте каждый интервал $\left(x_{i-1},x_i\right),\;x_{i-1}\notin I_k,x_i\notin I_k,\;2\le i\le k+p$ величине $R\left(i\right)$, называемой характеристикой интервала и вычисляемой по формуле

\begin{displaymath}
R\left(i\right)=rM\Delta_i+\frac{\left(z_i-z_{i-1}\right)^2}{rM\Delta_i}-2\left(z_i+z_{i-1}\right),
\end{displaymath}

где $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$ и $ r>1$ — параметр надежности метода.

\item  Выбрать интервал $\left[x_{t-1},x_t\right]$, которому соответствует максимальная характеристика, т.е.

\begin{displaymath}
R\left(t\right)=\max{\left\{R\left(i\right):\;x_{i-1}\notin I_k,x_i\notin I_k,\;2\le i\le k+p\right\}}.
\end{displaymath}

\item  Определить новую пробную точку $y^{k+p+1}=y\left(x^{k+p+1}\right)$, прообразом которой является $x^{k+p+1}\in\left(x_{t-1},x_t\right)$ по формуле

\begin{displaymath}
x^{k+p+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}\left(z_t-z_{t-1}\right)\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{M}\right]^N.
\end{displaymath}

\item 	Если $ k\ <\ 100\ast\ N$, то возвращаемся к шагу 1.
\item 	Если используем decision tree не впервые, то переходим к пункту 16.
\item 	Инициализируем decision tree (заполняем массив многомерных точек, массив значений функции, а так же создаем decision tree).
\item 	Тренируем дерево с помощью массивов полученных в предыдущем пункте.
\item 	Строим равномерную сетку, откладывая по каждой координате $\sqrt[N]{300}$ точек
\item 	Предсказываем значения функции в узлах сетки с помощью decision tree.
\item 	Находим ближайшую с точки зрения Евклидова расстояния точку к текущей.
\item 	Обходим соседей по описанному ранее принципу.
\item 	Если не нашлось значений функции меньших, чем в текущей точке, то запускаем метод Хука-Дживса.
\item 	Накапливаем массив из $100\ \ast\ N$ точек.
\item 	Все 100*N точек обрабатывает аналогично пункту 8-13, но без пункта 9 (без заполнения равномерной сетки, т.к. она у нас уже подготовлена).

\end{enumerate}

При вычислении очередной trial, главный процесс добавляет ее к множеству $I_k$ и отправляет рабочему процессу, который инициирует новую пробную точку в этой точке.
Главный процесс завершает алгоритм, если выполняется одно из двух условий: $\Delta_t<$ или $k+p>K_{max}$.
Вещественное число $\epsilon>0$ и целое число $K_{max}>0$ являются параметрами алгоритма и соответствуют точности поиска решения и максимальному количеству попыток соответственно.

Все полученные результаты были добавлены в систему globalizer \cite{fio_bib18}.
Если говорить более конкретно, то была реализована объединенная схема с запуском локальных методов в нужный момент, для чего была внедрена система  построения decision tree, равномерной сетки и прочих сопутствующих вычислений.

\begin{figure}[ht!]

	\begin{center}
		\begin{minipage}[h]{0.9\linewidth}
			\includegraphics[width=1\linewidth]{figure/fig3.png}
			\caption{Блок-схема объединения АГП и метода Хука-Дживса с использованием decision tree} %% подпись к рисунку
			\label{fig:fig3}
		\end{minipage}
	\end{center}
\end{figure}

\section{Эксперименты}

The numerical experiments were carried out on Lobachevsky supercomputer. Each supercomputer node was a two-processor one (two Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM). All methods considered in the above sections were implemented using Globalizer software \cite{globalizerSystem} (development language – C++). Intel C++ 17.0.0 compiler was used to assembly the system.

В работе \cite{fio_bib13, fio_bib17} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.

Ниже приведены результаты сравнения двух алгоритмов – стандартного АГП и его объединение с методом Хука-Дживса, при использовании decision tree. Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{fio_bib19}. Критерием остановки служило попадание trial в эпсилон окрестность истинного глобального минимума. 
При этом измерялось не только среднее число проводимых trial, но и среднее число trial.

В таблице \ref{tab:1} представлено среднее количество итераций, и ускорение по времени относительно последовательного запуска. Распараллеливание  осуществлялось с применением асинхронного MPI алгоритма. При этом, запуск производился на 4 процессах, а следовательно, на каждой итерации вычислялось 3 точки. Эксперименты производились с двумя алгоритмами: параллельный алгоритмом глобального поиска (АГПА) и объединением АГП и локального метода с использованием decision tree  на задачах разной размерности ($N$) и разного класса (Класс задачи). Значение в скобках (если оно есть) показывает количество не решенных задач. Это означает, что было достигнут максимум по числу проводимых итераций, но в окрестность глобального минимума точка так и не попала.

\begin{table}[h!]
    \caption{Среднее число итераций и среднее число trial, проводимое разными алгоритмами}
    \label{tab:1}
    \centering
    \begin{tabular}{|c|c|c|c|c|c|}
    \hline

        N & Класс задачи & \multicolumn{2}{c|}{Среднее число итераций} & \multicolumn{2}{c|}{Ускорение} \\ \hline
          & ~ & АГПА & decision tree & АГПА & decision tree \\ \hline
          & Simple & 2350.0 & 247.4 & 2348.0 & 390.0  \\ \hline
        2  & Hard & 4732.3 & 353.6 & 4730.3 & 1101.1  \\ \hline
          & Simple & 2115.5 & 579.5 & 2113.5 & 2441.4  \\ \hline
        3  & Hard & 5347.4 & 588.1 & 5345.4 & 2532.7  \\ \hline
          & Simple & 12168.9 & 768.2 & 12166.9 & 3582.2  \\ \hline
        4  & Hard & 25636.5 & 960.0 & 52634.5 & 5171.6  \\ \hline
          & Simple & 20633.5 & 978.0 & 20631.5 & 4840.4  \\ \hline
        5  & Hard & 161094.9 (4) & 1217.0 & 161092.9 (4) & 7115.5  \\ \hline
    \end{tabular}
\end{table}

Итак, ускорение по числу итераций неоспоримое, причем на задачах любой размерности. Среднее число trial тоже, практически везде, уменьшилось, но эти значения могут быть сильно выше, чем среднее число trial. Это связано с тем, что в trial учитываются и те точки, которые в процессе своей работы ставил локальный метод, а их может быть довольно много.

\section{Заключение}

Методы глобальной оптимизации являются областью активных научных исследований.  В результате работы удалось успешно объединить алгоритм глобального поиска вместе с поиском локального минимума методом Хука-Дживса. С целью экспериментального подтверждения теоретических свойств рассматриваемого алгоритма были так же проведены вычислительные эксперименты на серии из сотни тестовых задач.

При использовании в качестве определяющего правила о необходимости запуска локального метода decision tree, нам удалось получить достаточно хорошее ускорение. Применяя параллельную версию Алгоритма Глобального поиска ускорение удалось сохранить, что является несомненным преимуществом. Таким образом, наша схема позволяет получить преимущество от обоих из подходов (и от распараллеливания, и от запуска локального метода).

%
% ---- Bibliography ----
%
\bibliographystyle{spmpsci}
\begin{thebibliography}{6}
%

\bibitem{fio_bib1}
Ferreiro A.M., Garcia J.A., Lopez-Salas J.G., Vazquez C. An efficient implementation of parallel simulated annealing algorithm in GPUs // Journal of global optimization. — 2013. — Vol. 57, No. 3. — P. 863–890.

\bibitem{fio_bib2}
Garcia-Martinez J.M., Garzon E.M., Ortigosa P.M. A GPU implementation of a hybrid evolutionary algorithm: GPuEGO // Journal of super-computing. — 2014. — Vol. 70, No. 2. — P. 684–695.

\bibitem{fio_bib3}
Langdon W.B. Graphics processing units and genetic programming: an overview // Soft Computing. — 2011. — Vol. 15, No 8. — P. 1657–1669.

\bibitem{fio_bib4}
Евтушенко Ю.Г., Малкова В.У., Станевичюс А.А. Параллельный поиск глобального экстремума функций многих переменных // Ж. вычисл. матем. и матем. физ. — 2009. — Т. 49, №2. — С. 255–269.

\bibitem{fio_bib5}
He J., Verstak A., Watson L.T., Sosonkina M. Design and implementation of a mas-sively parallel version of DIRECT // Computational optimization and applications. — 2008. — Vol. 40, No. 2. — P. 217–245

\bibitem{fio_bib6}
Paulavicius R., Žilinskas J., Grothey A. Parallel branch and bound for global optimiza-tion with combination of Lipschitz bounds // Optimization methods and software. — 2011. — Vol. 26, No. 3. — P. 487–498.

\bibitem{fio_bib7}
Глобальная оптимизация: приложения и вычислительная сложность. %URL: http://hpc-education.unn.ru/ru/globopt/глобальная-оптимизация-приложения-и (дата обращения: 10.10.2020).

\bibitem{fio_bib8}
Захарова Е.М., Минашина И.К. Обзор методов многомерной оптимизации. Информационные процессы // том 14 - №3 – 2014 - С. 256-274.

\bibitem{fio_bib9}
Городецкий С. Ю. Методические материалы к лабораторной работе «Вычислительные методы поиска локальных минимумов функций», 2001.% URL: http://itmm.unn.ru/files/2016/09/MO_Lab2_LocOpt.pdf (дата обращения: 09.07.2022).

\bibitem{fio_bib10}
Шефов К. С., Степанова М. М. Реализация и применение параллельного алгоритма глобального поиска минимума к задаче оптимизации параметров молекулярно-динамического потенциала ReaxFF: %URL: http://crm.ics.org.ru/uploads/crmissues/crm_2015_3/15750.pdf, 2015 (дата обращения: 13.10.2020).

\bibitem{fio_bib11}
Gergel V.P. A global optimization algorithm for multivariate functions with lipschitzian first derivatives / Gergel V.P. // Journal of Global Optimization. – 1997. – Vol. 10, No. 3. – P. 257-281.

\bibitem{fio_bib12}
Barkalov K.A. A global optimization technique with an adaptive order of checking for constraints / Barkalov K.A., Strongin R.G. // Computational Mathematics and Mathematical Physics. – 2002. – Vol. 42, No. 9. – P. 1289-1300.

\bibitem{fio_bib13}
Gaviano, M. Software for generation of classes of test functions with known local and global minima for global optimization/ M. Gaviano, D. Lera, D. E. Kvasov, Y. D. Sergeyev // ACM Transactions on Mathematical Software. – 2003. – Vol. 29. – P. 469-480.

\bibitem{fio_bib14}
Химмельблау Д. Прикладное нелинейное программирование // М.: Мир - 1975 - 536 с.

\bibitem{fio_bib15}
Nelder J., Mead R. A simplex method for function minimization // Computer Journal - 7(4) – 1965 - P. 308-313.

\bibitem{fio_bib16}
OpenCV (Open Source Computer Vision) documentation. %URL: https://docs.opencv.org/4.x/dc/dd6/ml_intro.html (дата обращения: 10.07.2022)

\bibitem{fio_bib17}
Сергеев, Я.Д. Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. – 352 c.

\bibitem{fio_bib18}
Sysoyev, A.,  Barkalov, K.,  Sovrasov, V.,  Lebedev, I.,  Gergel, V. Globalizer – A parallel software system for solving global optimization problems. Lecture Notes in Computer Science. Volume 10421 LNCS, 2017, Pages 492-499

\bibitem{fio_bib19}
Sergeyev, Ya.D. Global search based on efficient diagonal partitions and a set of Lipschitz constants / Ya.D. Sergeyev, D.E. Kvasov // SIAM Journal on Optimization. – 2006. Vol. 16, No. 3. – P. 910–937

\bibitem{fio_bib20}
Стронгин Р.Г. Параллельные вычисления в задачах глобальной оптимизации / Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А. – М.: Издательство Московского университета, 2013. 280 с.

\end{thebibliography}
\end{document}

### 2023-01-15 2023/PAVT/ParallelDecisionTree.tex
Исправлен текст.

@@
 
-При поиске глобального минимума функции можно придерживаться различных алгоритмов: от основанных на идее случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3}, до детерминированных алгоритмов гарантирующих сходимость к глобальному минимуму \cite{fio_bib4, fio_bib5, fio_bib6}.
+При поиске глобального минимума функции можно придерживаться различных алгоритмов: от основанных на идее случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3}, до детерминированных алгоритмов, гарантирующих сходимость к глобальному минимуму \cite{fio_bib4, fio_bib5, fio_bib6}.
 Поскольку в реальных задачах глобальной оптимизации каждое вычисление значения функции представляет собой весьма трудоемкую задачу, необходимо уменьшить количество таких операций.
@@
 На этой идее основывается алгоритм глобального поиска (АГП) \cite{fio_bib7}.
-В данной работе мы попробуем объединить АГП и метод Хука-Дживса (метод локальной оптимизации) с помощью decision tree, с целью уменьшить количество проводимых итераций, а вместе с этим и количество вычислений целевой функции.
+В данной работе мы попробуем объединить АГП и метод локальной оптимизации (метод Хука-Дживса) с помощью decision tree, с целью уменьшить количество проводимых итераций, а вместе с этим и количество вычислений целевой функции.
 
@@
 Среди детерминированных различают методы нулевого порядка и градиентные (1-го и 2-го порядка) \cite{fio_bib8, fio_bib9}.
-В данной работе использовался метод Хука-Дживса.
 
@@
 Предсказание позволяет нам получить кусочно-линейную аппроксимацию по этому дереву.
-А дальше мы рассматриваем уже полученную по дереву аппроксимацию.
-Проверяем значения соседних точек к той, которую считаем подозрительной на минимум, и оцениваем попали мы в локальную ямку или нет.
-Причем, decision tree можно обобщить на многомерный случай (подробнее об этом расскажем дальше).
+После чего мы рассматриваем уже полученную по дереву аппроксимацию.
+Проверяем значения соседних точек к той, которую считаем подозрительной на минимум, и оцениваем попали мы в область притяжения локального минимума или нет.
 
@@
 
-В ННГУ им.
-Н.И.
-Лобачевского под руководством проф.
-Р.Г.
-Стронгина разработан эффективный подход к решению задач глобальной оптимизации.
-В рамках данного подхода решение многомерных задач сводится к решению серии вложенных задач меньшей размерности.
-Для эффективного решения быстро вычисляемых задач предлагается полностью перенести решение вложенной задачи на графический ускоритель.
-
-\section{Многомерный алгоритм глобального поиска}
+\section{Многомерный параллельный алгоритм глобального поиска}
 
@@
 \begin{equation}
-%\label{agp3_Ri}
+\label{agp3_Ri}
 R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1,
@@
 При решении тестовых задач, в которых точка глобального минимума $x^*$ является  известной, можно использовать также и критерий остановки по попаданию в окрестность глобального минимума.
-В этом случае метод прекращает работу, если если хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}) выполняется условие $\left|x_{t_j}-\ x^\ast\right| < \varepsilon.$
+В этом случае метод прекращает работу, если хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}) выполняется условие $\left|x_{t_j}-\ x^\ast\right| < \varepsilon.$
 
@@
 Decision tree — это метод автоматического анализа больших массивов данных, который применяется в машинном обучении.
-Оно представляет собой бинарное дерево (дерево, в котором каждый нелистовой узел имеет два дочерних узла).
+Decision tree представляет собой бинарное дерево (дерево, в котором каждый не листовой узел имеет два дочерних узла).
 Его можно использовать как для решения задач классификации, так и для регрессионных задач.
-В первом случае, каждый лист дерева помечается меткой класса, а несколько листьев могут иметь одну и ту же метку.
-Во втором же, каждому листу дерева присваивается константа, поэтому получаемая функция аппроксимации является кусочно-постоянной.
+При решении задач классификации, каждый лист дерева помечается меткой класса, при этом несколько листьев могут иметь одну и ту же метку.
+В случае построение регрессии, каждому листу дерева присваивается константа, как следствие  получаемая функция аппроксимации является кусочно-постоянной.
 
 Процедура прогнозирования по decision tree начинается с корня.
-От каждого нелистового узла процедура идет влево или вправо на основании значения некоторой переменной, индекс которой хранится в наблюдаемом узле.
+От каждого не листового узла процедура идет влево или вправо на основании значения некоторой переменной, индекс которой хранится в наблюдаемом узле.
 Возможны следующие переменные:
@@
 
-Здесь мы описали все лишь в общих чертах и не вдавались в подробности.
-Это связано с тем, что непосредственно сами decision tree мы не реализовывали, а лишь использовали готовые из библиотеки OpenCV (это библиотека алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения с открытым кодом).
+При реализации алгоритма  решения задач многомерной глобальной оптимизации с применением decision tree для выявления областей притяжения локальных минимумов нами были использованы  decision tree из библиотеки OpenCV (это библиотека алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения с открытым кодом).
 Подробнее с decision tree можно ознакомиться в \cite{fio_bib16}
 
-\section{Объединение метода Хука-Дживса и АГП в случае многомерных задач}
+\section{Объединение локального метода оптимизации и АГП для решения многомерных задач}
 
-Пришла пора описать как использовать decision tree в многомерном случае.
-В процессе работы АГП мы хотим понять, стоит ли использовать текущую точку, в качестве стартовой для локального метода или это будет бессмысленно.
-Для этого можно осмотреть соседние точки и если среди них нет ни одной такой, значение функции в которой меньше, чем в текущей, то можно предположить, что мы находимся в подобии локальной ямки и можно запустить локальный метод, чтобы быстрее сойтись к минимуму.
-Использование проекций многомерных точек на одномерный отрезок (то, что позволяют сделать кривые Пеано) нам для этих целей не подойдут сразу по нескольким причинам.
-Во-первых, вид функции после отображения очень сильно изменится и искать минимумы станет уже неэффективной идеей.
-Во-вторых, после отображения мы можем потерять информацию о взаимном расположении точек исходной функции (точки, расположенные близко в многомерном пространстве, могут оказаться на разных концах одномерного отрезка и наоборот).
+В текущем разделе приведем подробное описание того, как использовать decision tree в многомерном случае.
+В процессе работы АГП необходимо определить, стоит ли использовать текущую точку, в качестве стартовой для локального метода или нет.
+Для этого можно осмотреть соседние trial и если среди них нет ни одной такой, значение функции в которой меньше, чем в текущей, то можно предположить, что мы находимся в области притяжения локального минимума и можно запустить локальный метод, чтобы быстрее сойтись к минимуму.
+Использование проекций многомерных точек на одномерный отрезок после редукции размерности (то, что позволяют сделать кривые Пеано) нам для этих целей не подойдут по нескольким причинам.
+Во-первых, функция после отображения сильно изменяется, один локальный минимум в многомерном пространстве может разделиться на множество минимумов после отображения.
+Во-вторых, после отображения мы можем потерять информацию о взаимном расположении точек исходной функции.
+Точки, расположенные близко в многомерном пространстве, могут оказаться на разных концах одномерного отрезка и наоборот).
 
@@
 
-Поэтому мы будем работать именно с исходными точками.
-Но и тут все не так однозначно.
-Определить соседние точки в многомерном пространстве достаточно трудоемко, но тут нам как раз и помогут decision tree.
-Итак, перейдем к самому методу.
+Поэтому мы будем работать именно с исходными точками. Однако определить соседние точки в многомерном пространстве достаточно трудоемко, и для этого мы будем использовать decision tree.
 
 После проведения определенного числа trial (например, $100\ \ast\ N$), используем все накопившиеся точки для инициализации подходящей структуры данных и тренировки на ее основе decision tree.
-Чтобы проще было определить соседние точки, мы построим равномерную сетку с определенным шагом по каждой из размерностей.
-Но вычислять значение функции в каждой из точек полученной сетки было бы слишком трудозатратно, поэтому мы подадим их на вход предсказанию по дереву.
+Чтобы проще было определить соседние точки, мы построим равномерную сетку с определенным шагом по каждой из размерностей, после чего передадим эти точки на вход предсказанию по дереву.
 Теперь, имея равномерную сетку точек и зная значение функции в каждой из них, найдем ближайшую, с точки зрения евклидова расстояния, к исходной.
-Эта ближайшая точка – это, по сути, проекция изначальной точки на равномерную сетку и определить ее соседей уже не такая сложная задача (мы можем просто перебрать индексы)
+Эта ближайшая точка является проекцией изначальной точки на равномерную сетку и что бы определить  ее соседей, достаточно  просто перебрать индексы соседей.
+К тому же поскольку decision tree строит аппроксимацию исходной задачи (пусть и кусочно постоянную) мы можем оценить значение функции в тех областях, в которых trial ранее не проводились.
 
-При рассмотрении соседей необходимо учитывать следующее: если хоть один из соседей имеет значение функции меньшее, чем в текущей точке, то смысла запускать локальный метод нет; если какой-то из соседей имеет значение функции такое же, как в текущей, то его также необходимо проверить (найти всех его соседей и проверять их полностью аналогично); и только в том случае, если все значения функций соседей больше (включая соседей равных точек) мы можем запускать локальный метод.
+При рассмотрении соседей необходимо учитывать следующее: если хоть один из соседей имеет значение функции меньшее, чем в текущей точке, то запуск локального метода не нужен; если какой-то из соседей имеет значение функции такое же, как в текущей, то его также необходимо проверить (найти всех его соседей и проверять их аналогично); и только в том случае, если все значения функций соседей больше (включая соседей равных точек) мы можем запускать локальный метод.
 
-На рисунке  (\ref{fig:fig2}) отображены линии уровней оптимизационной задачи с отображением текущих точек trial, на рисунен  (\ref{fig:fig2_2}) отображено соответствующее decision tree.
+На рисунке  (\ref{fig:fig2}) отображены линии уровней оптимизационной задачи с отображением текущих  trial, и на рисунке  (\ref{fig:fig2_2}) отображено соответствующее decision tree.
 
@@
 
-Таким образом параллельный ассинхронный алгоритм глобальной оптимизации с использованием decision tree для выявления локальных областей состоит из следующих шагов:
+Таким образом параллельный асинхронный алгоритм глобальной оптимизации с использованием decision tree для выявления локальных областей состоит из следующих шагов:
 
@@
 \item 	Если $ k\ <\ 100\ast\ N$, то возвращаемся к шагу 1.
-\item 	Если используем decision tree не впервые, то переходим к пункту 16.
+\item 	Если используем decision tree не впервые, то переходим к пункту 15.
 \item 	Инициализируем decision tree (заполняем массив многомерных точек, массив значений функции, а так же создаем decision tree).
@@
 \item 	Обходим соседей по описанному ранее принципу.
-\item 	Если не нашлось значений функции меньших, чем в текущей точке, то запускаем метод Хука-Дживса.
+\item 	Если не нашлось значений функции меньших, чем в текущей точке, то запускаем локальный метод.
 \item 	Накапливаем массив из $100\ \ast\ N$ точек.
-\item 	Все 100*N точек обрабатывает аналогично пункту 8-13, но без пункта 9 (без заполнения равномерной сетки, т.к. она у нас уже подготовлена).
+\item 	Все 100*N точек обрабатывает аналогично пункту 8-14, но без пункта 9 (без заполнения равномерной сетки, т.к. она у нас уже подготовлена).
 
@@
 Все полученные результаты были добавлены в систему globalizer \cite{fio_bib18}.
-Если говорить более конкретно, то была реализована объединенная схема с запуском локальных методов в нужный момент, для чего была внедрена система  построения decision tree, равномерной сетки и прочих сопутствующих вычислений.
+Была реализована объединенная схема с запуском локальных методов при обнаружении областей притяжения локальных минимумов, для чего была внедрена система  использующая decision tree.
 
@@
 
-Ниже приведены результаты сравнения двух алгоритмов – стандартного АГП и его объединение с методом Хука-Дживса, при использовании decision tree.
+В таблице \ref{tab:1} отражено среднее число итераций , которые выполнил метод при решении серии задач из данных классов.
+Символ «>» отражает ситуацию, когда не все задачи класса были решены каким-либо методом.
+Количество нерешенных задач указано в скобках.
+
+\begin{table}[!ht]
+    \caption{Среднее число итераций}
+    \label{tab:1}
+    \centering
+    \begin{tabular}{|l|l|l|l|l|}
+    \hline
+        N & Problem class & DIRECT & DIRECTl & АГП  \\ \hline
+        4 & Simple & >47282 (4) & 18983 & 11953  \\ \hline
+        ~ & Hard & >95708 (7) & 68754 & 25263  \\ \hline
+        5 & Simple & >16057 (1) & 16758 & 15920  \\ \hline
+        ~ & Hard & >217215 (16) & >269064 (4) & >148342 (4)  \\ \hline
+    \end{tabular}
+\end{table}
+
+Как видно из таблицы 1, АГП превосходит методы DIRECT и DIRECTl на всех классах задач по среднему числу итераций.
+
+Ниже приведены результаты сравнения двух алгоритмов – асинхронного АГП и его объединение с локальным методом (методом Хука-Дживса) при использовании decision tree.
 Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{fio_bib19}.
 Критерием остановки служило попадание trial в эпсилон окрестность истинного глобального минимума.
-При этом измерялось не только среднее число проводимых trial, но и среднее число trial.
 
-В таблице \ref{tab:1} представлено среднее количество итераций, и ускорение по времени относительно последовательного запуска.
+В таблице \ref{tab:2} представлено среднее количество итераций, и ускорение по времени относительно последовательного запуска.
 Распараллеливание  осуществлялось с применением асинхронного MPI алгоритма.
 При этом, запуск производился на 4 процессах, а следовательно, на каждой итерации вычислялось 3 точки.
-Эксперименты производились с двумя алгоритмами: параллельный алгоритмом глобального поиска (АГПА) и объединением АГП и локального метода с использованием decision tree  на задачах разной размерности ($N$) и разного класса (Класс задачи).
 Значение в скобках (если оно есть) показывает количество не решенных задач.
@@
     \caption{Среднее число итераций и среднее число trial, проводимое разными алгоритмами}
-    \label{tab:1}
+    \label{tab:2}
     \centering
@@
           & Simple & 20633.5 & 978.0 & 20631.5 & 4840.4  \\ \hline
-        5  & Hard & 161094.9 (4) & 1217.0 & 161092.9 (4) & 7115.5  \\ \hline
+        5  & Hard & 161094.9 & 1217.0 & 161092.9 (4) & 7115.5  \\ \hline
     \end{tabular}
@@
 
-Итак, ускорение по числу итераций неоспоримое, причем на задачах любой размерности.
-Среднее число trial тоже, практически везде, уменьшилось, но эти значения могут быть сильно выше, чем среднее число trial.
-Это связано с тем, что в trial учитываются и те точки, которые в процессе своей работы ставил локальный метод, а их может быть довольно много.
+Как видно из таблицы ускорение по числу итераций достаточно велико, причем на задачах любой размерности.
 

### 2023-01-15 2023/PAVT/ParallelDecisionTree.tex
change abstract

@@
 В работе рассматривается решение задач многомерной глобальной оптимизации с применением decision tree для выявления областей притяжения локальных минимумов.
-Искомая функция задана как черный ящик и может быть выдвинуто лишь общее предположение о том, что она удовлетворяет условию Липшица с неизвестной константой.
-В ходе решения задачи алгоритмом глобального поиска производится анализ позволяющий принять решение о запуске локального метода.
-Локальные методы играют роль локального уточнения и призваны ускорить сходимость алгоритма.
-Для подтверждения теории, были приведены вычислительных экспериментов, демонстрирующие ускорение при решении серии тестовых задач.
-\dots
+Целевая функция задачи задана ''как черный ящик''.
+Мы лишь предполагаем, что она удовлетворяет условию Липшица с неизвестной константой. 
+Для поиска глобального минимума в задачах такого типа используется алгоритм глобального поиска.
+В настоящей работе мы предлагаем способ выделения окрестностей локальных экстремумов целевой функции на основе анализа накопленной поисковой информации.
+Проведение такого анализа средствами машинного обучения позволяет принять решение о запуске локального метода, который может ускорить сходимость алгоритма.
+Выдвинутое предположение подтверждается результатами вычислительных экспериментов, демонстрирующих ускорение при решении серии тестовых задач.
 % We would like to encourage you to list your keywords within
@@
 
-При поиске глобального минимума функции можно придерживаться различных алгоритмов: от основанных на идее случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3}, до детерминированных алгоритмов, гарантирующих сходимость к глобальному минимуму \cite{fio_bib4, fio_bib5, fio_bib6}.
-Поскольку в реальных задачах глобальной оптимизации каждое вычисление значения функции представляет собой весьма трудоемкую задачу, необходимо уменьшить количество таких операций.
-Этого можно добиться целенаправленным выбором вариантов в процессе поиска оптимального решения.
-На этой идее основывается алгоритм глобального поиска (АГП) \cite{fio_bib7}.
-В данной работе мы попробуем объединить АГП и метод локальной оптимизации (метод Хука-Дживса) с помощью decision tree, с целью уменьшить количество проводимых итераций, а вместе с этим и количество вычислений целевой функции.
+Для решения указанной задачи известен широкий спектр алгоритмов, он основаных на идее случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3} метаэвристических алгоритмов, до детерминированных алгоритмов, гарантирующих сходимость к глобальному минимуму \cite{fio_bib4, fio_bib5, fio_bib6}.
+Поскольку в реальных задачах глобальной оптимизации каждое вычисление значения функции (называемое в дальнейшем trial) представляет собой весьма трудоемкую операцию, необходимо уменьшить количество таких trials.
+Этого можно добиться целенаправленным выбором вариантов в процессе поиска оптимального решения, отсекая бесперспективные подобласти поиска и исследуя только те, в которых может содержаться решение задачи.
+На этой идее основывается global search algorithm (GSA) \cite{fio_bib7}.
+В данной работе мы попробуем объединить АГП и метод локальной оптимизации (метод Хука-Дживса) с целью уменьшить количество проводимых испытаний.
+При этом принятие решения о запуске локального метода будет осуществляться с помощью decision tree.
 
-Алгоритмы поиска локального экстремума предназначены для определения лишь одного из локальных экстремумов на множестве допустимых решений, в котором целевая функция принимает максимальное или минимальное значение.
-При их построении могут использоваться как детерминированный спуск в область экстремума, так и случайный поиск.
+Алгоритмы поиска локального экстремума предназначены для определения лишь одного из локальных экстремумов на множестве допустимых решений, в котором целевая функция принимает экстремальное (максимальное или минимальное) значение.
+%При их построении могут использоваться как детерминированный спуск в область экстремума, так и случайный поиск. 
 Среди детерминированных различают методы нулевого порядка и градиентные (1-го и 2-го порядка) \cite{fio_bib8, fio_bib9}.

### 2023-01-16 2023/PAVT/ParallelDecisionTree.tex
first section

@@
 
-Алгоритмы поиска локального экстремума предназначены для определения лишь одного из локальных экстремумов на множестве допустимых решений, в котором целевая функция принимает экстремальное (максимальное или минимальное) значение.
+Алгоритмы локальной оптмизации предназначены для определения лишь одного из локальных экстремумов на множестве допустимых решений, в котором целевая функция принимает экстремальное (максимальное или минимальное) значение.
 %При их построении могут использоваться как детерминированный спуск в область экстремума, так и случайный поиск. 
-Среди детерминированных различают методы нулевого порядка и градиентные (1-го и 2-го порядка) \cite{fio_bib8, fio_bib9}.
+Среди локальных методов традиционно выделяют методы нулевого порядка и градиентные методы  (1-го и 2-го порядков) \cite{fio_bib8, fio_bib9}.
+Отметим, что применение методов первого порядка (не говоря уже о втором) в задачах с функциями вида ''черный ящик'' затруднено, т.к.
+здесь возникает проблема численной оценки градиента.
+Поэтому мы будем рассматривать в дальнейшем только методы нулевого порядка.
 
-Для того, чтобы понять, в какой момент лучше всего запускать локальный метод, мы используем алгоритм на основе decision tree.
-Мы накапливаем определенное число точек и лишь когда их станет достаточно начинаем обработку.
-Эти полученные точки мы используем для тренировки decision tree и предсказания по нему.
-Предсказание позволяет нам получить кусочно-линейную аппроксимацию по этому дереву.
-После чего мы рассматриваем уже полученную по дереву аппроксимацию.
-Проверяем значения соседних точек к той, которую считаем подозрительной на минимум, и оцениваем попали мы в область притяжения локального минимума или нет.
+Для того, чтобы понять, в какой момент лучше всего запускать локальный метод в рамках глобального поиска, мы будем использовать алгоритм на основе decision tree.
+На первой стадии поиска происходит накоплении поисковой информации в виде результатов проведенных испытаний.
+Результаты испытаний мы используем для тренировки decision tree, что позволяет получить кусочно-линейную аппроксимацию целевой функции и  предсказывать на ее основе поведение целевой функции.
+Для этого бы будем сравнивать значения функции в точках, соседних к той, которую считаем подозрительной на локальный минимум.
+И на основе кусочной-линейной аппроксимации принимать решение, находится очередная точка испытания в области притяжения локального минимума или нет.
+
+%Структура статьи
 
@@
 
-Рассмотрим задачу поиска глобального минимума функции $\varphi(y)$ в гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1\le\ i\le\ n \}$ При этом будем предполагать, что функция удовлетворяет условию Липшица с априори неизвестной константой $L$.
-Что соответствует ограниченности изменения значений функции при ограниченной вариации аргумента.
+Рассмотрим задачу поиска глобального минимума функции $\varphi(y)$ в гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1\le\ i\le\ n \}$ При этом будем предполагать, что функция удовлетворяет условию Липшица с априори неизвестной константой $L$, что соответствует ограниченности изменения значений функции при ограниченной вариации аргумента.
 Это предположение можно интерпретировать (применительно к прикладным задачам) как отражение ограниченности мощностей, порождающих изменения в моделируемой системе.
@@
 \end{equation}
-где $a,b \in R$ --- заданные векторы.
+где $a,b \in R$ есть заданные векторы.
 
@@
 
-Численное решение задачи (\ref{sec:problem})   сводится к построению оценки $ y_k^\ast\in\ D$ , отвечающей некоторому понятию близости к точке $y^\ast$ (например, ${||y^\ast-y}_k^\ast||\le\ \varepsilon$, где $\varepsilon\geq0$ есть заданная точность) на основе конечного числа $k$ вычислений значений оптимизируемой функции.
-Относительно класса рассматриваемых задач предполагается, что оптимизируемая функция $\varphi(y)$ может быть задана не аналитически, в виде формулы, а алгоритмически, как результат работы некоторой подпрограммы или библиотеки.
+Численное решение задачи (\ref{sec:problem})   сводится к построению оценки $ y_k^\ast\in\ D$ , отвечающей некоторому понятию близости к точке $y^\ast$ (например, ${||y^\ast-y}_k^\ast||\le\ \varepsilon$, где $\varepsilon\geq0$ есть заданная точность) на основе конечного числа $k$ вычислений значений целевой функции.
+Относительно класса рассматриваемых задач предполагается, что оптимизируемая функция $\varphi(y)$ может быть задана алгоритмически, как результат работы некоторой подпрограммы или библиотеки.
 
@@
 
-Основная идея данного подхода заключается в том, что минимизируемая функция рассматривается как реализация некоторого случайного процесса.
+Основная идея данного подхода заключается в том, что целевая функция рассматривается как реализация некоторого случайного процесса.
 Решающие правила алгоритма конструируются таким образом, что очередная итерация проводится в точке глобального минимума математического ожидания значений функции.
 Эта точка записывается в список известных значений и итерации повторяются.
-Так происходит до тех пор, пока не достигнут один из выбранных критериев остановки: расстояние между точками отрезка не становится меньше заданного значения или новые точки не попадают в окрестность истинного глобального минимума.
-Еще одним вариантом остановки работы алгоритма является достижение установленного максимума возможных итераций.
+Так происходит до тех пор, пока не достигнут один из выбранных критериев остановки: расстояние между точками проведенных испытаний становится меньше заданного значения или же достигается установленное максимальное число итераций.
 \cite{fio_bib10}

### 2023-01-17 2023/PAVT/ParallelDecisionTree.tex
a lot of fixes in Russian

@@
 %При их построении могут использоваться как детерминированный спуск в область экстремума, так и случайный поиск. 
-Среди локальных методов традиционно выделяют методы нулевого порядка и градиентные методы  (1-го и 2-го порядков) \cite{fio_bib8, fio_bib9}.
+Среди локальных методов традиционно выделяют методы нулевого порядка и градиентные методы  (1-го и 2-го порядков) \cite{fio_bib8,fio_bib9}.
 Отметим, что применение методов первого порядка (не говоря уже о втором) в задачах с функциями вида ''черный ящик'' затруднено, т.к.
@@
 
-Рассмотрим задачу поиска глобального минимума функции $\varphi(y)$ в гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1\le\ i\le\ n \}$ При этом будем предполагать, что функция удовлетворяет условию Липшица с априори неизвестной константой $L$, что соответствует ограниченности изменения значений функции при ограниченной вариации аргумента.
+Рассмотрим задачу поиска глобального минимума функции $\varphi(y)$ в гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1\le\ i\le\ n \}$.
+При этом будем предполагать, что функция удовлетворяет the Lipschitz condition с априори неизвестной константой $L$, что соответствует ограниченности изменения значений функции при ограниченной вариации аргумента.
 Это предположение можно интерпретировать (применительно к прикладным задачам) как отражение ограниченности мощностей, порождающих изменения в моделируемой системе.
@@
 
-где функция $\varphi(y(x^\ast))$  удовлетворяют уже равномерному условию Гельдера
+где функция $\varphi(y(x^\ast))$  удовлетворяют более общему H{\"o}lder condition
 
 \begin{displaymath}
-\left|\varphi (y \left(x_1\right))- \varphi (y \left(x_2\right)\right )|\le\ H\left|x_1-x_2\right|^\frac{1}{N},\ x_1,\ x_2\epsilon[0,1]		
+\left|\varphi (y \left(x_1\right))- \varphi (y \left(x_2\right)\right )|\le\ H\left|x_1-x_2\right|^\frac{1}{N},\ x_1,\ x_2\epsilon[0,1].
 \end{displaymath}
 
-Поэтому, не ограничивая общности, можно рассматривать минимизацию одномерной функции $f(x)\ =\ \varphi(y(x)), \ x\epsilon [0,1]$, удовлетворяющей условию Гельдера.
+Поэтому вместо исходной задачи минимизации функции $\varphi(y)$ в области $D$ можно рассматривать минимизацию одномерной функции $f(x)\ =\ \varphi(y(x))$, удовлетворяющей the H{\"o}lder condition, при $ x\in [0,1]$.
 
-Схема алгоритма:
+Основные шаги parallel global search algorithm состоят в следующем.
 
@@
 
-Если выполнено $n\geq1$ итераций, которым соответствуют $k=k(n)$ проведенных trial in point $x^i, 1\leq i\leq k$, то trial $x^{k+1},\ldots,x^{k+p}$  следующей $(n+1)$-й итерации будут определять следующим образом.
+Если выполнено $n \geq 1$ итераций, которым соответствуют $k=k(n)$ проведенных испытаний в точках $x^i, 1\leq i\leq k$, то точки $x^{k+1},\ldots,x^{k+p}$ поисковых испытаний следующей $(n+1)$-й итерации будут вычисляться в результате выполнения следующих действий.
 
@@
 
-\item  Перенумеровать (нижним индексом) trial ранее проведенных  $x^i, 1\leq i\leq k$, а также граничные точки отрезка [0,1] в порядке возрастания координаты:
+\item  Переупорядочить (нижним индексом) точки $x^i, 1\leq i\leq k$, а также граничные точки отрезка [0,1] в порядке возрастания координаты:
  \begin{equation}
@@
 
-\item  Вычислить текущие нижние оценки $M$ неизвестной константы Гельдера $H$:
+\item  Вычислить текущие нижние оценки $M$ неизвестной H{\"o}lder constani $H$:
  \begin{equation}
@@
 	\end{equation}
-где $r>1$ --- заданный параметр алгоритма.
+где $r>1$ есть параметр алгоритма.
+Данный параметр регулирует надежность алгоритма: большие значения $r$ обеспечивают гарантированное нахождение глобального минимума, выбор меньшего значения --- ускоряет сходимость алгоритма
 
@@
 \end{equation}	
-и выбрать $p$ интервалов с номерами $t_j,\ 1\le\ j\le\ p$, с наибольшими значениями характеристики.
+и выбрать $p$ интервалов с номерами $t_j,\ 1\le\ j\le\ p$, с наибольшими значениями характеристик.
 
@@
 
-Очередные trial проводятся параллельно, в вычисленных по формулам (\ref{agp5_x1}), (\ref{agp4_xi}) точках $x^{k+j},\ 1\leq j\leq p$.
+Очередные $p$ испытаний проводятся параллельно, в вычисленных по формулам (\ref{agp5_x1}), (\ref{agp4_xi}) точках $x^{k+j},\ 1\leq j\leq p$.
+После проведения испытаний их результаты заносятся в информационную базу, и алгоритм переходит к расчету новых точек испытаний.
 
-Отметим, что в прикладных оптимизационных задачах процесс проведения trial является, как правило, значительно более трудоемким по сравнению с работой вычислительных правил алгоритма.
+Отметим, что в прикладных оптимизационных задачах процесс проведения trial является, как правило, значительно более трудоемким по сравнению с расчетом точки испытания.
 
 Алгоритм останавливает свою работу в случае, если хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}) выполняется условие \(\Delta_{t_j} < \varepsilon\).
-Данный критерий остановки (наряду с обычным для итерационных методов критерием, ограничивающим число выполненных итераций) используется в задачах оптимизации, в которых точка глобального минимума $x^*$ заранее неизвестна.
+Данный критерий остановки (наряду с обычным для итерационных методов критерием, ограничивающим число выполненных итераций) используется в прикладных задачах оптимизации, в которых точка глобального минимума $x^*$ заранее неизвестна.
 
@@
 
-Метод Хука–Дживса – это комбинация исследующего поиска по направлениям и поиска по образцу \cite{fio_bib14, fio_bib15}.
+Метод Хука–Дживса относится к классу методов нулевого порядка.
+Его вычислительные правила представляет собой комбинацию исследующего поиска (для выбора направления) и поиска по выбранному направлению \cite{fio_bib14, fio_bib15}.
 
-Исследующий поиск определяется следующим образом: 
+Исследующий поиск выполняется следующим образом: 
 \begin{enumerate}
-\item	Определяется значение шага (оно различно для каждого направления координат и может измениться во время поиска). 
-\item	Шаг поиска считается успешным, если значение целевой функции в контрольной точке не превышает значение исходной целевой функции. 
+\item	Определяется значение шага (оно различно для каждой координаты и может измениться во время поиска). 
+\item	Шаг поиска считается успешным, если значение целевой функции в контрольной точке не превышает значение целевой функции в исходной точке. 
 \item	В противном случае нужно вернуться к предыдущему пункту и сделать шаг в обратном направлении. 
-\item	После перебора всех N координат исследовательский поиск заканчивается. Полученная точка называется базовой точкой.
+\item	После перебора всех $N$ координат исследующий поиск заканчивается. Полученная точка называется базовой точкой.
 \end{enumerate}
 
-Что касается поиска по образцу, то он заключается в реализации шага из найденной в ходе исследующего поиска базовой точки вдоль прямой, соединяющей ее с предыдущей базовой точкой.
+Что касается поиска по направлению, то он заключается в реализации шага из найденной в ходе исследующего поиска базовой точки вдоль прямой, соединяющей ее с предыдущей базовой точкой.
 Величина этого шага определяется задаваемым заранее параметром.
@@
 
-Decision tree — это метод автоматического анализа больших массивов данных, который применяется в машинном обучении.
+Дерево принятия решений (decision tree) --- это средство, используемое для автоматического анализа больших массивов данных, который применяется в машинном обучении.
 Decision tree представляет собой бинарное дерево (дерево, в котором каждый не листовой узел имеет два дочерних узла).
-Его можно использовать как для решения задач классификации, так и для регрессионных задач.
-При решении задач классификации, каждый лист дерева помечается меткой класса, при этом несколько листьев могут иметь одну и ту же метку.
-В случае построение регрессии, каждому листу дерева присваивается константа, как следствие  получаемая функция аппроксимации является кусочно-постоянной.
+Его можно использовать как для решения задач классификации, так и для задач регрессии.
+При решении задач классификации каждый лист дерева помечается меткой класса, при этом несколько листьев могут иметь одну и ту же метку.
+В случае построения регрессии каждому листу дерева присваивается константа, поэтому получаемая функция аппроксимации является кусочно-постоянной.
 
-Процедура прогнозирования по decision tree начинается с корня.
-От каждого не листового узла процедура идет влево или вправо на основании значения некоторой переменной, индекс которой хранится в наблюдаемом узле.
+%Я предлагаю следующий абзац удалить
+%Процедура прогнозирования по decision tree начинается с корня.
+От каждого не листового узла процедура идет влево или вправо на основании значения некоторой переменной, индекс которой хранится в текущем узле.
 Возможны следующие переменные:
-Упорядоченные переменные, значения которых сравнивается с порогом (который также хранится в узле). Если значение меньше порога, процедура идет влево. В противном случае - вправо.
-Категориальные переменные - значение дискретной переменной проверяется на предмет того, принадлежит ли оно определенному подмножеству значений (также хранящемуся в узле) из набора допустимых величин.
+%Упорядоченные переменные, значения которых сравнивается с порогом (который также хранится в узле). Если значение меньше порога, процедура идет влево. В противном случае - вправо.
+%Категориальные переменные - значение дискретной переменной проверяется на предмет того, принадлежит ли оно определенному подмножеству значений (также хранящемуся в узле) из набора допустимых величин.
 Если принадлежит, то идем влево.
 В противном случае - вправо.
+%Как только алгоритм достигает конечного узла, в качестве ответа процедуры прогнозирования выбирается значение, присвоенное этому узлу.
 
-Как только алгоритм достигает конечного узла, в качестве ответа процедуры прогнозирования выбирается значение, присвоенное этому узлу.
-
-При реализации алгоритма  решения задач многомерной глобальной оптимизации с применением decision tree для выявления областей притяжения локальных минимумов нами были использованы  decision tree из библиотеки OpenCV (это библиотека алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения с открытым кодом).
+При реализации алгоритма  решения задач многомерной глобальной оптимизации с применением decision tree для выявления областей притяжения локальных минимумов нами были использованы алгоритмы из библиотеки OpenCV.
+OpenCV --- библиотека алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения с открытым кодом.
 Подробнее с decision tree можно ознакомиться в \cite{fio_bib16}
 
-\section{Объединение локального метода оптимизации и АГП для решения многомерных задач}
+\section{Объединение локального метода оптимизации и GSA для решения многомерных задач}
 
-В текущем разделе приведем подробное описание того, как использовать decision tree в многомерном случае.
-В процессе работы АГП необходимо определить, стоит ли использовать текущую точку, в качестве стартовой для локального метода или нет.
-Для этого можно осмотреть соседние trial и если среди них нет ни одной такой, значение функции в которой меньше, чем в текущей, то можно предположить, что мы находимся в области притяжения локального минимума и можно запустить локальный метод, чтобы быстрее сойтись к минимуму.
-Использование проекций многомерных точек на одномерный отрезок после редукции размерности (то, что позволяют сделать кривые Пеано) нам для этих целей не подойдут по нескольким причинам.
-Во-первых, функция после отображения сильно изменяется, один локальный минимум в многомерном пространстве может разделиться на множество минимумов после отображения.
-Во-вторых, после отображения мы можем потерять информацию о взаимном расположении точек исходной функции.
-Точки, расположенные близко в многомерном пространстве, могут оказаться на разных концах одномерного отрезка и наоборот).
+В текущем разделе приведем подробное описание того, как мы использовали decision tree для выявления областей притяжения локальных экстремумов. 
+В процессе работы GSA необходимо определить, стоит ли использовать текущую точку в качестве стартовой для локального метода или нет.
+Для этого можно проверить точки соседних испытаний и если среди них нет ни одной такой, значение функции в которой меньше, чем в текущей, то можно предположить, что мы находимся в области притяжения локального минимума.
+В этом случае из текущей точки можно запустить локальный метод, который быстро сойдется к локальному минимуму функции.
+Сделать это можно только в многомерном пространстве, а не на одномерном отрезке после редукции размерности.  
+Во-первых, редуцированная функция $\varphi(y(x))$ меняет свои свойства: один локальный минимум в многомерном пространстве может разделиться на множество минимумов после отображения.
+Во-вторых, после отображения мы можем потерять информацию о взаимном расположении точек в исходном пространстве.
+Точки, расположенные близко в многомерном пространстве, могут оказаться существенно разнесенными на одномерном отрезке.
 
@@
 
-Поэтому мы будем работать именно с исходными точками. Однако определить соседние точки в многомерном пространстве достаточно трудоемко, и для этого мы будем использовать decision tree.
+Поэтому определять области притяжения локальных экстремумов мы будем в исходном многомерном пространстве.
+Однако определить соседние точки в многомерном пространстве достаточно трудоемко, и для этого мы будем использовать decision tree.
 
 После проведения определенного числа trial (например, $100\ \ast\ N$), используем все накопившиеся точки для инициализации подходящей структуры данных и тренировки на ее основе decision tree.
-Чтобы проще было определить соседние точки, мы построим равномерную сетку с определенным шагом по каждой из размерностей, после чего передадим эти точки на вход предсказанию по дереву.
+Чтобы проще было определить соседние точки, мы построим равномерную сетку с определенным шагом по каждой из координат, после чего вычислим значения аппроксимации целевой функции в этих точках с использованием дерева решений.
 Теперь, имея равномерную сетку точек и зная значение функции в каждой из них, найдем ближайшую, с точки зрения евклидова расстояния, к исходной.
-Эта ближайшая точка является проекцией изначальной точки на равномерную сетку и что бы определить  ее соседей, достаточно  просто перебрать индексы соседей.
-К тому же поскольку decision tree строит аппроксимацию исходной задачи (пусть и кусочно постоянную) мы можем оценить значение функции в тех областях, в которых trial ранее не проводились.
+Эта ближайшая точка является проекцией исходной точки на равномерную сетку, на которой можно легко определить ее соседей.
+А так как decision tree строит аппроксимацию исходной функции (пусть и кусочно постоянную), то мы можем оценить значение функции в тех областях, в которых испытания ранее не проводились.
 
-При рассмотрении соседей необходимо учитывать следующее: если хоть один из соседей имеет значение функции меньшее, чем в текущей точке, то запуск локального метода не нужен; если какой-то из соседей имеет значение функции такое же, как в текущей, то его также необходимо проверить (найти всех его соседей и проверять их аналогично); и только в том случае, если все значения функций соседей больше (включая соседей равных точек) мы можем запускать локальный метод.
+При рассмотрении соседей необходимо учитывать следующее: если хоть один из соседей имеет значение функции меньшее, чем в текущей точке, то запуск локального метода не нужен. Е
+Если хотя бы в одной из соседних точек значение функции такое же, как в текущей, то ее также необходимо проверить (т.е.
+найти всех ее соседей и проверять их аналогично).
+И только в том случае, если все значения функции в соседних точках больше, мы можем запускать локальный метод.
 
-На рисунке  (\ref{fig:fig2}) отображены линии уровней оптимизационной задачи с отображением текущих  trial, и на рисунке  (\ref{fig:fig2_2}) отображено соответствующее decision tree.
+На рисунке  (\ref{fig:fig2}) отображены линии уровня целевой функции с отображением точек испытаний, а на рисунке  (\ref{fig:fig2_2}) отображено соответствующее decision tree.
 
@@
 
-Идея асинхронной схемы в том, чтобы не дожидаться окончания вычисления всей отправленной группы точек, а отправлять новую точку, как только была закончена работа с первой из старых.
+Способ организации параллельных вычислений, описанный в предыдущем разделе, подразумевает синхронное проведение $p$ поисковых испытаний в точках, вычисленных на текущей итерации.
+При этом переход к следующей итерации будет выполнен лишь после окончания проведения всех испытаний.
+В случае различной трудоемкости проведения испытаний в разных точках области поиска данный подход может привести к дисбалансу вычислительной нагрузки. 
+Исправить этот недостаток можно введением асинхронности.
+Идея асинхронной схемы состоит в том, чтобы не дожидаться окончания проведения всех $p$ испытаний, а инициировать проведение нового испытания сразу же, как только было закончено хотя бы  одно их них.
 
 Предположим, что главный процесс вычисляет одну точку следующего trial на каждой итерации и отправляет ее рабочему процессу для выполнения trial.
-В то же время выполнение trial рабочим процессом – гораздо более затратная в вычислительном отношении операция, чем выбор мастером новой trial, исключающей простаивающие рабочие процессы.
+Отметим, что в прикладных оптимизационных задачах время выполнение trial рабочим процессом -- гораздо более затратная в вычислительном отношении операция, чем выбор мастером точки нового trial.
+Тем самым простой рабочих процессов будет пренебрежимо мал.
 В этом случае (в отличие от синхронных параллельных алгоритмов) общее количество trial, выполняемых каждым рабочим процессом, будет зависеть от вычислительных затрат на выполнение конкретного trial и не может быть оценено заранее.
@@
 
- В начале поиска главный процесс (допустим, это процесс $0$) инициирует параллельное выполнение $p$ trial в $p$ разных точках области поиска.
+ В начале поиска главный процесс (допустим, это процесс с идентификатором $0$) инициирует параллельное выполнение $p$ trial в $p$ разных точках области поиска.
 
-Две из этих точек являются граничными, а остальные – внутренними, т. е. в точках $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$, где
-$x^1=,x^p=,x^i\in\left(0,1\right),i=2,\ldots,p$
+Две из этих точек являются граничными, а остальные -- внутренними, т.е. в точках $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$, где
+$x^1=,x^p=,x^i\in\left(0,1\right),i=2,\ldots,p$.
 
@@
 
-Каждый рабочий процесс, завершив в какой-то момент вычисление функции, отправляет главному процессу значение.
-В свою очередь, главный процесс выбирает новую пробную точку $x^{k+p+1}$ для рабочего процесса в соответствии с описанными ниже правилами.
-Заметим, что в этом случае мы будем иметь набор прообразов пробных точек
+Каждый рабочий процесс, завершив в какой-то момент вычисление функции, отправляет главному процессу вычисленное значение.
+В свою очередь, главный процесс выбирает новую точку $x^{k+p+1}$ для рабочего процесса в соответствии с описанными ниже правилами.
+Заметим, что в этом случае мы будем иметь набор preimages of the trial points
 $I_k=\left\{x^{k+1},x^{k+2},\ldots,x^{k+p}\right\}$
-на котором trial уже начаты, но еще не завершены.
+в которых испытания уже начаты, но еще не завершены.
 
@@
 \begin{enumerate}
-\item  Перенумеровать множество прообразов точек trial
+\item  Упорядочить нижним индексом (в порядке возрастания) множество preimages of the trial points
 
@@
 
-содержащие все прообразы, на которых либо завершились, либо идут trial в порядке возрастания (по нижнему индексу) так, что
+содержащие все точки, в которых либо завершились, либо идут trial, т.е. получить упорядоченный набор
 
@@
 
-\item  Вычислите значения
+\item  Вычислить значения
 
@@
 Значения $z_i$ в точках $x_i\in I_k$ не определены, так как trial в точках $x_i\in I_k$ еще не завершены.
-Если значение $M$ равно $0$, то установите $M=1$.
+Если значение $M$ равно $0$, то установить $M=1$.
 
-\item  Сопоставьте каждый интервал $\left(x_{i-1},x_i\right),\;x_{i-1}\notin I_k,x_i\notin I_k,\;2\le i\le k+p$ величине $R\left(i\right)$, называемой характеристикой интервала и вычисляемой по формуле
+\item  Сопоставить каждыму интервалу $\left(x_{i-1},x_i\right),\;x_{i-1}\notin I_k,x_i\notin I_k,\;2\le i\le k+p$ его характеристику $R\left(i\right)$, вычисляемую по формуле
 
@@
 
-где $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$ и $ r>1$ — параметр надежности метода.
+где $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$ и $ r>1$ -- параметр надежности метода.
 
@@
 
-\item  Определить новую пробную точку $y^{k+p+1}=y\left(x^{k+p+1}\right)$, прообразом которой является $x^{k+p+1}\in\left(x_{t-1},x_t\right)$ по формуле
+\item  Определить новую точку испытания $y^{k+p+1}=y\left(x^{k+p+1}\right)$, прообразом которой является $x^{k+p+1}\in\left(x_{t-1},x_t\right)$, по формуле
 
@@
 
-\item 	Если $ k\ <\ 100\ast\ N$, то возвращаемся к шагу 1.
-\item 	Если используем decision tree не впервые, то переходим к пункту 15.
-\item 	Инициализируем decision tree (заполняем массив многомерных точек, массив значений функции, а так же создаем decision tree).
+\item 	Если $ k\ <\ 100\ast\ N$, то вернуться к шагу 1.
+
+\item 	Если используем decision tree не впервые, то перейти к пункту 15.
+
+\item 	Инициализируем decision tree (заполняем массив многомерных точек, массив значений функции, а так же создаем decision tree). %объединить
 \item 	Тренируем дерево с помощью массивов полученных в предыдущем пункте.
-\item 	Строим равномерную сетку, откладывая по каждой координате $\sqrt[N]{300}$ точек
+
+\item 	Строим равномерную сетку, откладывая по каждой координате $\sqrt[N]{300}$ точек %объединить
 \item 	Предсказываем значения функции в узлах сетки с помощью decision tree.
-\item 	Находим ближайшую с точки зрения Евклидова расстояния точку к текущей.
+
+\item 	Находим ближайшую с точки зрения Евклидова расстояния точку к текущей. %объединить
 \item 	Обходим соседей по описанному ранее принципу.
+
 \item 	Если не нашлось значений функции меньших, чем в текущей точке, то запускаем локальный метод.
-\item 	Накапливаем массив из $100\ \ast\ N$ точек.
+
+\item 	Накапливаем массив из $100\ \ast\ N$ точек. %это не понятно....
 \item 	Все 100*N точек обрабатывает аналогично пункту 8-14, но без пункта 9 (без заполнения равномерной сетки, т.к. она у нас уже подготовлена).
@@
 
-При вычислении очередной trial, главный процесс добавляет ее к множеству $I_k$ и отправляет рабочему процессу, который инициирует новую пробную точку в этой точке.
+При вычислении точки очередного испытания главный процесс добавляет ее к множеству $I_k$ и отправляет рабочему процессу, который инициирует проведение испытания в ней.
 Главный процесс завершает алгоритм, если выполняется одно из двух условий: $\Delta_t<$ или $k+p>K_{max}$.
-Вещественное число $\epsilon>0$ и целое число $K_{max}>0$ являются параметрами алгоритма и соответствуют точности поиска решения и максимальному количеству попыток соответственно.
-
-Все полученные результаты были добавлены в систему globalizer \cite{fio_bib18}.
-Была реализована объединенная схема с запуском локальных методов при обнаружении областей притяжения локальных минимумов, для чего была внедрена система  использующая decision tree.
+Вещественное число $\epsilon>0$ и целое число $K_{max}>0$ являются параметрами алгоритма и соответствуют точности поиска решения и максимальному количеству испытаний соответственно.
 

### 2023-01-17 2023/PAVT/ParallelDecisionTree.tex
Поправлено описание алгоритма

@@
 В случае построения регрессии каждому листу дерева присваивается константа, поэтому получаемая функция аппроксимации является кусочно-постоянной.
+В нашем случае decision tree строит функцию $\varphi(y)$ кусочно постоянную аппроксимацию$\varphi(y)$  в многомерном пространстве, обозначим как $z' = \psi(y)$ значение вычисленное по decision tree в точке $y$.
 
@@
 
+\item  Получить вычисленное  значение функции с процесса $j$, добавить новый trial $z_j = f(y(x_j))$, к множеству $V$.
+
+%X_k=\left\{x^1,x^2,\ldots,x^{k+p}\right\}
+
 \item 	Если $ k\ <\ 100\ast\ N$, то вернуться к шагу 1.
 
-\item 	Если используем decision tree не впервые, то перейти к пункту 15.
+%\item 	Если используем decision tree не впервые, то перейти к пункту 15.
 
-\item 	Инициализируем decision tree (заполняем массив многомерных точек, массив значений функции, а так же создаем decision tree). %объединить
-\item 	Тренируем дерево с помощью массивов полученных в предыдущем пункте.
+\item Создаем decision tree по множеству $I_k$, получаем функцию аппроксимации  $\psi(y)$.
 
-\item 	Строим равномерную сетку, откладывая по каждой координате $\sqrt[N]{300}$ точек %объединить
-\item 	Предсказываем значения функции в узлах сетки с помощью decision tree.
+\item 	Если используем decision tree впервые, то строим равномерную сетку
 
-\item 	Находим ближайшую с точки зрения Евклидова расстояния точку к текущей. %объединить
-\item 	Обходим соседей по описанному ранее принципу.
+\begin{displaymath}
+Y'=\{ y'\in\ R^N:\ a_i\le\  {y'_i}^k \le\ b_i,\ 1\le\ i\le\ N,\ 1\le k\le\ \sqrt[N]{300}  \}
+\end{displaymath}
 
-\item 	Если не нашлось значений функции меньших, чем в текущей точке, то запускаем локальный метод.
+\item 	Вычисляем значения аппроксимации: $Z' = \{ z'=  \psi(y'), y' \in Y'\}$
 
-\item 	Накапливаем массив из $100\ \ast\ N$ точек. %это не понятно....
-\item 	Все 100*N точек обрабатывает аналогично пункту 8-14, но без пункта 9 (без заполнения равномерной сетки, т.к. она у нас уже подготовлена).
+\item Для всех точек $y'\in V$:
+
+Находим точку $y'_q$ ближайшую  к $y'$,
+
+Производим обход соседей $y'_q$ по описанному ранее принципу.
+
+Если ни у одной соседней точки не нашлось значений функции меньших, чем в $y'_q$, то запускаем локальный метод.
+
+Очищаем множество $V$.
 

### 2023-01-18 2023/PAVT/ParallelDecisionTree.tex
before translation of the first part

@@
 %
-\title{Combining Global and Local Optimization Algorithms Using Decision Trees}
+\title{Об использовании деревьев решений для идентификации локальных экстремумов в параллельном алгоритме глобальной оптимизации}
 %
-\titlerunning{Global Optimization Using Decision Trees}  % abbreviated title (for running head)
+\titlerunning{Global Optimization ...}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 В работе рассматривается решение задач многомерной глобальной оптимизации с применением decision tree для выявления областей притяжения локальных минимумов.
-Целевая функция задачи задана ''как черный ящик''.
+Целевая функция задачи задана как ''черный ящик''.
 Мы лишь предполагаем, что она удовлетворяет условию Липшица с неизвестной константой. 
@@
 
-Для решения указанной задачи известен широкий спектр алгоритмов, он основаных на идее случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3} метаэвристических алгоритмов, до детерминированных алгоритмов, гарантирующих сходимость к глобальному минимуму \cite{fio_bib4, fio_bib5, fio_bib6}.
+В работе рассматриваются параллельные алгоритмы для решения задач многомерной глобальной оптимизации.
+Такие задачи часто возникают в случае, когда требуется выбрать значения параметров исследуемой математической модели, при которых результаты моделирования наилучшим образом соответствуют экспериментальным данным.
+Для решения задач указанного класса известен широкий спектр алгоритмов, он основанных на идее случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3} метаэвристических алгоритмов, до детерминированных алгоритмов, гарантирующих сходимость к глобальному минимуму \cite{fio_bib4, fio_bib5, fio_bib6}.
 Поскольку в реальных задачах глобальной оптимизации каждое вычисление значения функции (называемое в дальнейшем trial) представляет собой весьма трудоемкую операцию, необходимо уменьшить количество таких trials.
@@
 На этой идее основывается global search algorithm (GSA) \cite{fio_bib7}.
-В данной работе мы попробуем объединить АГП и метод локальной оптимизации (метод Хука-Дживса) с целью уменьшить количество проводимых испытаний.
+В данной работе мы попробуем объединить GSA и метод локальной оптимизации (Hooke--Jeeves method) с целью уменьшить количество проводимых испытаний.
 При этом принятие решения о запуске локального метода будет осуществляться с помощью decision tree.
 
-Алгоритмы локальной оптмизации предназначены для определения лишь одного из локальных экстремумов на множестве допустимых решений, в котором целевая функция принимает экстремальное (максимальное или минимальное) значение.
-%При их построении могут использоваться как детерминированный спуск в область экстремума, так и случайный поиск. 
+Алгоритмы локальной оптимизации предназначены для определения лишь одного из локальных экстремумов на множестве допустимых решений, в котором целевая функция принимает экстремальное (максимальное или минимальное) значение.
 Среди локальных методов традиционно выделяют методы нулевого порядка и градиентные методы  (1-го и 2-го порядков) \cite{fio_bib8,fio_bib9}.
@@
 
-%Структура статьи
+Основная часть статьи имеет следующую структуру. 
+В Section \ref{SecPS} рассмотрена постановка задачи глобальной оптимизации, сделаны основные предположения о целевой функции .
+В Section \ref{SecA} приведено описание используемых методов и подходов.
+В частности, в Section \ref{SecGSA} обсуждается основная идея global search algorithm, приведены его вычислителные правила.
+Section \ref{SecHG} содержит описание используемого метода локального поиска.
+Способ построения аппроксимаций функции с использованием деревьев решений описан в Section \ref{SecDT}.
+Новый алгоритм, объединяющий глобальный и локальный поиск на основе использования деревьев решений, приведен в Section \ref{SecGSAL}.
+В следующем разделе обсуждаются особенности асинхронной схемы распараллеливания нового алгоритма.
+Section \ref{SecR} содержит результаты экспериментов, проведенных на параллельной вычислительной системе с distributed memory.
+В Заключении представлены основные выводы по проделанной работе.
 
-\section{Постановка задачи}
+\
+\section{Problem statement}\label{SecPS}
 
@@
 Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (сетки) в области параметров, и выборе наилучшего значения функции на данной сетке.
-Вычислительные затраты на решение задачи растут экспоненциально с ростом размерности.
+При использовании равномерных сеток вычислительные затраты на решение задачи растут экспоненциально с ростом размерности.
 
-\section{Многомерный параллельный алгоритм глобального поиска}
+\section{Используемые алгоритмы}\label{SecA}
 
-Основная идея данного подхода заключается в том, что целевая функция рассматривается как реализация некоторого случайного процесса.
-Решающие правила алгоритма конструируются таким образом, что очередная итерация проводится в точке глобального минимума математического ожидания значений функции.
+Основная идея подхода к построению более экономной неравномерной сетки заключается в том, что ее точки вычисляются последовательно, а целевая функция рассматривается как реализация некоторого случайного процесса.
+Решающие правила алгоритма построения сетки конструируются таким образом, что очередная точка сетки соответствует точке глобального минимума математического ожидания значений функции.
 Эта точка записывается в список известных значений и итерации повторяются.
@@
 Поэтому вместо исходной задачи минимизации функции $\varphi(y)$ в области $D$ можно рассматривать минимизацию одномерной функции $f(x)\ =\ \varphi(y(x))$, удовлетворяющей the H{\"o}lder condition, при $ x\in [0,1]$.
+
+\subsection{Многомерный параллельный алгоритм глобального поиска}\label{SecGSA}
 
@@
 
-\section{Метод Хука-Дживса}
+\subsection{Hooke--Jeeves method}\label{SecHG}
 
@@
 
-\section{Decision tree}
+\subsection{Decision tree}\label{SecDT}
 
@@
 
-%Я предлагаю следующий абзац удалить
-%Процедура прогнозирования по decision tree начинается с корня.
-От каждого не листового узла процедура идет влево или вправо на основании значения некоторой переменной, индекс которой хранится в текущем узле.
-Возможны следующие переменные:
-%Упорядоченные переменные, значения которых сравнивается с порогом (который также хранится в узле). Если значение меньше порога, процедура идет влево. В противном случае - вправо.
-%Категориальные переменные - значение дискретной переменной проверяется на предмет того, принадлежит ли оно определенному подмножеству значений (также хранящемуся в узле) из набора допустимых величин.
-Если принадлежит, то идем влево.
-В противном случае - вправо.
-%Как только алгоритм достигает конечного узла, в качестве ответа процедуры прогнозирования выбирается значение, присвоенное этому узлу.
-
 При реализации алгоритма  решения задач многомерной глобальной оптимизации с применением decision tree для выявления областей притяжения локальных минимумов нами были использованы алгоритмы из библиотеки OpenCV.
@@
 
-\section{Объединение локального метода оптимизации и GSA для решения многомерных задач}
+\section{Объединение локального метода оптимизации и GSA для решения многомерных задач}\label{SecGSAL}
 
@@
 Сделать это можно только в многомерном пространстве, а не на одномерном отрезке после редукции размерности.  
-Во-первых, редуцированная функция $\varphi(y(x))$ меняет свои свойства: один локальный минимум в многомерном пространстве может разделиться на множество минимумов после отображения.
-Во-вторых, после отображения мы можем потерять информацию о взаимном расположении точек в исходном пространстве.
+Во-первых, редуцированная функция $\varphi(y(x))$ меняет свои свойства: один локальный минимум в многомерном пространстве может разделиться на множество минимумов после редукции размерности.
+Во-вторых, после использования отображения $y(x)$ мы можем потерять информацию о взаимном расположении точек в исходном пространстве.
 Точки, расположенные близко в многомерном пространстве, могут оказаться существенно разнесенными на одномерном отрезке.
@@
 	\begin{center}
-		\begin{minipage}[h]{0.8\linewidth}
+		\begin{minipage}[h]{0.6\linewidth}
 			\includegraphics[width=1\linewidth]{figure/fig2.png}
@@
 
-При рассмотрении соседей необходимо учитывать следующее: если хоть один из соседей имеет значение функции меньшее, чем в текущей точке, то запуск локального метода не нужен. Е
+При рассмотрении соседей необходимо учитывать следующее: если хоть один из соседей имеет значение функции меньшее, чем в текущей точке, то запуск локального метода не нужен.
 Если хотя бы в одной из соседних точек значение функции такое же, как в текущей, то ее также необходимо проверить (т.е.
@@
 
-На рисунке  (\ref{fig:fig2}) отображены линии уровня целевой функции с отображением точек испытаний, а на рисунке  (\ref{fig:fig2_2}) отображено соответствующее decision tree.
+На Fig. \ref{fig:fig2} отображены линии уровня целевой функции с отображением точек испытаний, а на Fig. \ref{fig:fig2_2} отображено соответствующее decision tree.
 
-\section{Асинхронная схема распараллеливания с использованием MPI }
+\section{Асинхронная схема распараллеливания алгоритма}\label{SecASP}
 
@@
 
-\section{Эксперименты}
+% переводить до этого раздела
+\section{Результаты экспериментов}\label{SecR}
 
@@
 
-\section{Заключение}
+\section{Заключение}\label{SecC}
 

### 2023-01-20 2023/problem-info/artricle.tex
Расширенная версия статьи по деревьям решений.

\documentclass[12pt, a4paper, russian]{article}
\usepackage[margin=2cm]{geometry}
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{amsmath}
\usepackage{subfig}
\usepackage{caption}
\usepackage[compress]{cite}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{titling}
\usepackage[hidelinks]{hyperref}
\usepackage{setspace}
\singlespacing

% To prevent LaTeX from hyphenating
\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\makeatletter \renewcommand*{\ALG@name}{Алгоритм} \makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%
%%%%%% Информация на английском языке
%%%%%%

\title{Global optimization algorithm using decision trees to find local extrema \thanks{The work was supported by the oneAPI Center of Excellence, by the Ministry of Science and Higher Education of the Russian Federation (project no. 0729-2020-0055), and by the Research and Education Mathematical Center (project no.  075-02-2021-1394).}}

\author{D.~I.~Silenko, I.~G.~Lebedev\\
	\small{Lobachevsky State University of Nizhny Novgorod, 603022, Nizhny Novgorod, Russia}
}
\date{}
\maketitle

\begin{small}
\textbf{Abstract:} The paper considers the solution of multidimensional global optimization problems using decision trees to identify areas of attraction of local minima. The local method is used for local refinement and is designed to speed up the convergence of the algorithm. The desired function satisfies the Lipschitz condition with an unknown constant. To confirm the theory, computational experiments were presented demonstrating acceleration when solving a series of test problems.

\textbf{Key words:} Global optimization, multiextremal functions, parallel computing, decision tree.
\end{small}

% References
\renewcommand{\refname}{References}
\bibliographystyle{IEEEtran}
\bibliography{articleEng}

%%%%%% 
%%%%%% Информация на русском языке
%%%%%%
\emptythanks
\title{Алгоритм глобальной оптимизации, использующий деревья решений для выявления локальных экстремумов\thanks{Работа выполнена при финансовой поддержке Министерства науки и высшего образования РФ (проект № 0729-2020-0055) и научно-образовательного математического центра “Математика технологий будущего” (соглашение № 075-02-2021-1394).}}

\author{Д. И. Силенко, И. Г. Лебедев\\
	\small{Нижегородский государственный университет им. Н. И. Лобачевского, 603022, Н.Новгород, Россия} 
}
\date{}
\maketitle

УДК 004.222+004.272.25

\vspace{\baselineskip}

\begin{small}
\textbf{Аннотация:} В работе рассматривается решение задач многомерной глобальной оптимизации с применением деревьев решений для выявления областей притяжения локальных минимумов. Локальный метод используется для локального уточнения и призван ускорить сходимость алгоритма. Искомая функция удовлетворяет условию Липшица с неизвестной константой. Для подтверждения теории были приведены вычислительные эксперименты, демонстрирующие ускорение при решении серии тестовых задач.

\textbf{Ключевые слова:} глобальная оптимизация, локальная оптимизация, многоэкстремальные функции, деревья решений.
\end{small}

\section{Введение}

При поиске глобального минимума функции можно придерживаться различных алгоритмов: от основанных на идее случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3}, до детерминированных алгоритмов, гарантирующих сходимость к глобальному минимуму \cite{fio_bib4, fio_bib5, fio_bib6}. Поскольку в реальных задачах глобальной оптимизации каждое вычисление значения функции представляет собой весьма трудоемкую задачу, необходимо уменьшить количество таких операций. Этого можно добиться целенаправленным выбором вариантов в процессе поиска оптимального решения. На этой идее основывается алгоритм глобального поиска (АГП). В данной работе мы попробуем объединить АГП и метод локальной оптимизации (метод Хука-Дживса) с помощью деревьев решений, с целью уменьшить количество проводимых итераций, а вместе с этим и количество вычислений целевой функции.

Алгоритмы поиска локального экстремума предназначены для определения лишь одного из локальных экстремумов на множестве допустимых решений, в котором целевая функция принимает максимальное или минимальное значение. При их построении могут использоваться как детерминированный спуск в область экстремума, так и случайный поиск. Среди детерминированных различают методы нулевого порядка и градиентные (1-го и 2-го порядка).

Для того, чтобы понять, в какой момент лучше всего запускать локальный метод, мы используем алгоритм на основе деревьев решений. Мы накапливаем определенное число точек и лишь когда их станет достаточно начинаем обработку. Эти полученные точки мы используем для тренировки деревьев решений и предсказания по нему. Предсказание позволяет нам получить кусочно-линейную аппроксимацию по этому дереву. После чего мы рассматриваем уже полученную по дереву аппроксимацию. Проверяем значения соседних точек к той, которую считаем подозрительной на минимум, и оцениваем попали мы в область притяжения локального минимума или нет.

\section{Постановка задачи}

Рассмотрим задачу поиска глобального минимума функции $\varphi(y)$ в гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1\le\ i\le\ n \}$ При этом будем предполагать, что функция удовлетворяет условию Липшица с априори неизвестной константой $L$. Что соответствует ограниченности изменения значений функции при ограниченной вариации аргумента. Это предположение можно интерпретировать (применительно к прикладным задачам) как отражение ограниченности мощностей, порождающих изменения в моделируемой системе.

\begin{equation} \label{sec:problem}   
\varphi(y^*) = min\{\varphi(y):y\in D\}, D = \{y \in R^N : a_i \leq y_i \leq b_i, 1 \leq i \leq N \},
\end{equation}
где $a,b \in R$ --- заданные векторы.

\begin{displaymath}
|\varphi(y_1)-\varphi(y_2)|\leq L\parallel y_1-y_2 \parallel
,y_1,y_2 \in D, 0<L< \infty.
\end{displaymath}

Численное решение задачи (\ref{sec:problem})   сводится к построению оценки $ y_k^\ast\in\ D$ , отвечающей некоторому понятию близости к точке $y^\ast$ (например, ${||y^\ast-y}_k^\ast||\le\ \varepsilon$, где $\varepsilon\geq0$ есть заданная точность) на основе конечного числа $k$ вычислений значений оптимизируемой функции. Относительно класса рассматриваемых задач предполагается, что оптимизируемая функция $\varphi(y)$ может быть задана не аналитически, в виде формулы, а алгоритмически, как результат работы некоторой подпрограммы или библиотеки.

Задачи многоэкстремальной оптимизации имеют существенно более высокую трудоемкость решения по сравнению с другими типами оптимизационных задач, т.к. глобальный оптимум является интегральной характеристикой решаемой задачи и требует исследования всей области поиска. Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (сетки) в области параметров, и выборе наилучшего значения функции на данной сетке. Вычислительные затраты на решение задачи растут экспоненциально с ростом размерности.

\section{Многомерный алгоритм глобального поиска}

Основная идея данного подхода заключается в том, что минимизируемая функция рассматривается как реализация некоторого случайного процесса. Решающие правила алгоритма конструируются таким образом, что очередная итерация проводится в точке глобального минимума математического ожидания значений функции. Эта точка записывается в список известных значений и итерации повторяются. Так происходит до тех пор, пока не достигнут один из выбранных критериев остановки: расстояние между точками отрезка не становится меньше заданного значения или новые точки не попадают в окрестность истинного глобального минимума. Еще одним вариантом остановки работы алгоритма является достижение установленного максимума возможных итераций. \cite{fio_bib10}

При решении многомерных задач применяется редукция размерности (т.е. сведение многомерной задачи к эквивалентной одномерной) с использованием кривых Пеано. Они позволяют свести многомерную задачу минимизации в области $D$  к одномерной задаче минимизации на отрезке $[0, 1]$

\begin{equation*}
\phi({y(x}^\ast))\ =\ min\{\phi(y(x)):\ x\epsilon[0,\ 1]\}
\end{equation*}
где функция  удовлетворяют уже равномерному условию Гельдера
\begin{equation*}
\left|\phi (y \left(x_1\right))- \phi (y \left(x_2\right)\right )|\le\ H\left|x_1-x_2\right|^\frac{1}{N},\ x_1,\ x_2\epsilon[0,1]		
\end{equation*}

Поэтому, не ограничивая общности, можно рассматривать минимизацию одномерной функции $f(x)\ =\ \phi(y(x)), \ x\epsilon [0,1]$, удовлетворяющей условию Гельдера.

Схема алгоритма:

На подготовительном шаге параллельно проводятся $p$ поисковых испытаний в произвольных внутренних точках $x^1, ...,x^p$ отрезка $[0,1]$, что соответствует первой итерации  алгоритма.

Если выполнено $n\geq1$ итераций, которым соответствуют $k=k(n)$ проведенных поисковых испытаний в точках $x^i, 1\leq i\leq k$, то точки $x^{k+1},\ldots,x^{k+p}$ испытаний следующей $(n+1)$-й итерации будут определять следующим образом.

\begin{enumerate}

\item  Перенумеровать (нижним индексом) точки ранее проведенных испытаний $x^i, 1\leq i\leq k$, а также граничные точки отрезка [0,1] в порядке возрастания координаты:
 \begin{equation}
\label{agp1_sort}
	0=x_0<\ x_1<\ ...\ <x_{k+1}=1.
	\end{equation}
	и сопоставить им значения $z_i=f(x_i)$.

\item  Вычислить текущие нижние оценки $M$ неизвестной константы Гельдера $H$:
 \begin{equation}
\label{agp2_mu}
	\mu=max\left\{\frac{|z_i-z_{i-1}|}{{{(x}_i-x_{i-1})}^{1/N}},\ i=1,\ldots,k\right\},\ M=\ \left\{\begin{matrix}r\mu,\ \mu>0,\\1,\ \mu=0,\\\end{matrix}\right.\
	\end{equation}
где $r>1$ --- заданный параметр алгоритма.

\item  Для каждого интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить величину $R(i)$, называемую \textit{характеристикой} интервала, в соответствии с формулами
\begin{equation}
\label{agp3_R1}
R(1)=2\Delta_1-4\dfrac{z_1}{M}, \; R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
\end{equation}
\begin{equation}
\label{agp3_Ri}
R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1,
\end{equation}
где \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).

\item   Упорядочить характеристики $R\left(i\right),\ 1\leq i \leq k+1,$ в порядке невозрастания 
\begin{equation}
\label{agp4_R_sort}
	R\left(t_1\right)\geq\ R\left(t_2\right)\geq...\geq\ R\left(t_k\right)\geq\ R(t_{k+1}),\ 
\end{equation}	
и выбрать интервал с номерами $t$, с наибольшими значениями характеристики.

\item  В выбранном интервале вычислить точку $x^{k+1}$, в соответствии с формулами
\begin{equation}
\label{agp5_x1}
	x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\ t_j=1,\ t_j=k+1,
\end{equation}	
	\begin{equation}
\label{agp4_xi}	
	x^{k+1}=\frac{x_{t_j}+x_{t_j-1}}{2}-sign\left(z_{t_j}-z_{t_j-1}\right)\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N,\ 1<t_j<k+1.
\end{equation}

\end{enumerate}

Отметим, что в прикладных оптимизационных задачах процесс проведения испытания является, как правило, значительно более трудоемким по сравнению с работой вычислительных правил алгоритма.

Алгоритм останавливает свою работу в случае, если значения $t$ из (\ref{agp4_R_sort}) выполняется условие \(\Delta_{t_j} < \varepsilon\). Данный критерий остановки (наряду с обычным для итерационных методов критерием, ограничивающим число выполненных итераций) используется в задачах оптимизации, в которых точка глобального минимума $x^*$ заранее неизвестна.

При решении тестовых задач, в которых точка глобального минимума $x^*$ является  известной, можно использовать также и критерий остановки по попаданию в окрестность глобального минимума. В этом случае метод прекращает работу, если если хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}) выполняется условие $\left|x_{t_j}-\ x^\ast\right| < \varepsilon.$

В качестве итоговой оценки глобально-оптимального решения рассматриваемой задачи выбираются значения 
\begin{equation}
f_k^*=\min_{1\leq i \leq k}f(x_i), \; x_k^*=arg \min_{1\leq i \leq k}f(x_i).
\end{equation}

Обоснование данного способа организации вычислений см. в \cite{fio_bib20}. Модификации, учитывающие наличие ограничений-неравенств в задаче, а также информация о производной целевой функции, представлены в \cite{fio_bib12, fio_bib9, fio_bib11}.

\section{Метод Хука-Дживса}

Метод Хука–Дживса --- это комбинация исследующего поиска по направлениям и поиска по образцу \cite{fio_bib14, fio_bib15}.

Исследующий поиск определяется следующим образом: 
\begin{enumerate}
\item	Определяется значение шага (оно различно для каждого направления координат и может измениться во время поиска). 
\item	Шаг поиска считается успешным, если значение целевой функции в контрольной точке не превышает значение исходной целевой функции. 
\item	В противном случае нужно вернуться к предыдущему пункту и сделать шаг в обратном направлении. 
\item	После перебора всех N координат исследовательский поиск заканчивается. Полученная точка называется базовой точкой.
\end{enumerate}

Что касается поиска по образцу, то он заключается в реализации шага из найденной в ходе исследующего поиска базовой точки вдоль прямой, соединяющей ее с предыдущей базовой точкой. Величина этого шага определяется задаваемым заранее параметром.

\begin{figure}[!h]
	\begin{center}
		\begin{minipage}[h]{0.8\linewidth}
			\includegraphics[width=1\linewidth]{figure/fig1.png}
			\caption{Пример итераций алгоритма Хука-Дживса} %% подпись к рисунку
			\label{fig:fig1}
		\end{minipage}
	\end{center}
\end{figure}

\section{Деревья решений}

Деревья решений --- это метод автоматического анализа больших массивов данных, который применяется в машинном обучении. Деревья решений представляет собой бинарное дерево (дерево, в котором каждый не листовой узел имеет два дочерних узла). Его можно использовать как для решения задач классификации, так и для регрессионных задач. При решении задач классификации, каждый лист дерева помечается меткой класса, при этом несколько листьев могут иметь одну и ту же метку. В случае построение регрессии, каждому листу дерева присваивается константа, как следствие  получаемая функция аппроксимации является кусочно-постоянной.

Процедура прогнозирования по дереву решения начинается с корня. От каждого не листового узла процедура идет влево или вправо на основании значения некоторой переменной, индекс которой хранится в наблюдаемом узле. Возможны следующие переменные:
Упорядоченные переменные, значения которых сравнивается с порогом (который также хранится в узле). Если значение меньше порога, процедура идет влево. В противном случае --- вправо.
Категориальные переменные --- значение дискретной переменной проверяется на предмет того, принадлежит ли оно определенному подмножеству значений (также хранящемуся в узле) из набора допустимых величин. Если принадлежит, то идем влево. В противном случае --- вправо. 
Как только алгоритм достигает конечного узла, в качестве ответа процедуры прогнозирования выбирается значение, присвоенное этому узлу.

В нашем случае дерево решений строит функцию $\varphi(y)$ кусочно постоянную аппроксимацию$\varphi(y)$  в многомерном пространстве, обозначим как $z' = \psi(y)$ значение вычисленное по decision tree в точке $y$.

При реализации алгоритма  решения задач многомерной глобальной оптимизации с применением деревьев решений для выявления областей притяжения локальных минимумов нами были использованы  деревья решений из библиотеки OpenCV (это библиотека алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения с открытым кодом). Подробнее с  деревьями решений можно ознакомиться в \cite{fio_bib16}

\section{Объединение локального метода и АГП в случае многомерных задач}

В текущем разделе приведем подробное описание того, как использовать деревьев решений в многомерном случае. В процессе работы АГП необходимо определить, стоит ли использовать текущую точку, в качестве стартовой для локального метода или нет. Для этого можно осмотреть соседние точки испытания и если среди них нет ни одной такой, значение функции в которой меньше, чем в текущей, то можно предположить, что мы находимся в области притяжения локального минимума и можно запустить локальный метод, чтобы быстрее сойтись к минимуму. Использование проекций многомерных точек на одномерный отрезок после редукции размерности (то, что позволяют сделать кривые Пеано) нам для этих целей не подойдут по нескольким причинам. Во-первых, функция после отображения сильно изменяется, один локальный минимум в многомерном пространстве может разделиться на множество минимумов после отображения. Во-вторых, после отображения мы можем потерять информацию о взаимном расположении точек исходной функции. Точки, расположенные близко в многомерном пространстве, могут оказаться на разных концах одномерного отрезка и наоборот).

%\begin{figure}[ht!]
%	\begin{center}
%		\begin{minipage}[h]{0.8\linewidth}
%			\includegraphics[width=1\linewidth]{figure/fig2.png}
%			\caption{Функция из генератора gkls до использования кривых Пеано} %% подпись к рисунку
%			\label{fig:fig2}
%		\end{minipage}
%	\end{center}
%\end{figure}

%\begin{figure}[ht!]
%	\begin{center}
%		\begin{minipage}[h]{0.8\linewidth}
%			\includegraphics[width=1\linewidth]{figure/fig2.png}
%			\caption{Рисунок с двумерным деревом решений} %% подпись к рисунку
%			\label{fig:fig2_2}
%		\end{minipage}
%	\end{center}
%\end{figure}

Поэтому мы будем работать именно с исходными точками. Однако определить соседние точки в многомерном пространстве достаточно трудоемко, и для этого мы будем использовать деревья решений.

После проведения определенного числа испытаний (например, $100\ \ast\ N$), используем все накопившиеся точки для инициализации подходящей структуры данных и тренировки на ее основе дерева решений. Чтобы проще было определить соседние точки, мы построим равномерную сетку с определенным шагом по каждой из размерностей, после чего передадим эти точки на вход предсказанию по дереву. Теперь, имея равномерную сетку точек и зная значение функции в каждой из них, найдем ближайшую, с точки зрения евклидова расстояния, к исходной. Эта ближайшая точка является проекцией изначальной точки на равномерную сетку и что бы определить  ее соседей, достаточно  просто перебрать индексы соседей. К тому же поскольку дерево решений строит аппроксимацию исходной задачи (пусть и кусочно постоянную) мы можем оценить значение функции в тех областях, в которых испытания ранее не проводились.

При рассмотрении соседей необходимо учитывать следующее: если хоть один из соседей имеет значение функции меньшее, чем в текущей точке, то запуск локального метода не нужен; если какой-то из соседей имеет значение функции такое же, как в текущей, то его также необходимо проверить (найти всех его соседей и проверять их аналогично); и только в том случае, если все значения функций соседей больше (включая соседей равных точек) мы можем запускать локальный метод. Модифицированная версия алгоритма глобального поиска выглядит следующим образом:

\begin{enumerate}

	\item  Перенумеровать (нижним индексом) точки ранее проведенных испытаний $x^i, 1\leq i\leq k$, а также граничные точки отрезка [0,1] в порядке возрастания координаты:
	\begin{equation}
		\label{agp1_sort}
		0=x_0<\ x_1<\ ...\ <x_{k+1}=1.
	\end{equation}
	и сопоставить им значения $z_i=f(x_i)$.

	\item  Вычислить текущие нижние оценки $M$ неизвестной константы Гельдера $H$:
	\begin{equation}
		\label{agp2_mu}
		\mu=max\left\{\frac{|z_i-z_{i-1}|}{{{(x}_i-x_{i-1})}^{1/N}},\ i=1,\ldots,k\right\},\ M=\ \left\{\begin{matrix}r\mu,\ \mu>0,\\1,\ \mu=0,\\\end{matrix}\right.\
	\end{equation}
	где $r>1$ --- заданный параметр алгоритма.

	\item  Для каждого интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить величину $R(i)$, называемую \textit{характеристикой} интервала, в соответствии с формулами
	\begin{equation}
		\label{agp3_R1}
		R(1)=2\Delta_1-4\dfrac{z_1}{M}, \; R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
	\end{equation}
	\begin{equation}
		\label{agp3_Ri}
		R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1,
	\end{equation}
	где \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).

	\item   Упорядочить характеристики $R\left(i\right),\ 1\leq i \leq k+1,$ в порядке невозрастания 
	\begin{equation}
		\label{agp4_R_sort}
		R\left(t_1\right)\geq\ R\left(t_2\right)\geq...\geq\ R\left(t_k\right)\geq\ R(t_{k+1}),\ 
	\end{equation}	
	и выбрать интервал с номерами $t$, с наибольшими значениями характеристики.

	\item  В выбранном интервале вычислить точку $x^{k+1}$, в соответствии с формулами
	\begin{equation}
		\label{agp5_x1}
		x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\ t_j=1,\ t_j=k+1,
	\end{equation}	
	\begin{equation}
		\label{agp4_xi}	
		x^{k+1}=\frac{x_{t_j}+x_{t_j-1}}{2}-sign\left(z_{t_j}-z_{t_j-1}\right)\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N,\ 1<t_j<k+1.
	\end{equation}

\item  Получить вычисленное  значение функции с процесса $j$, добавить новый trial $z_j = f(y(x_j))$, к множеству $V$.

%X_k=\left\{x^1,x^2,\ldots,x^{k+p}\right\}

\item 	Если $ k\ <\ 100\ast\ N$, то вернуться к шагу 1.

%\item 	Если используем decision tree не впервые, то перейти к пункту 15.

\item Создаем decision tree по множеству $I_k$, получаем функцию аппроксимации  $\psi(y)$.

\item 	Если используем decision tree впервые, то строим равномерную сетку

\begin{displaymath}
	Y'=\{ y'\in\ R^N:\ a_i\le\  {y'_i}^k \le\ b_i,\ 1\le\ i\le\ N,\ 1\le k\le\ \sqrt[N]{300}  \}
\end{displaymath}

\item 	Вычисляем значения аппроксимации: $Z' = \{ z'=  \psi(y'), y' \in Y'\}$

\item Для всех точек $y'\in V$:

Находим точку $y'_q$ ближайшую  к $y'$,

Производим обход соседей $y'_q$ по описанному ранее принципу.

Если ни у одной соседней точки не нашлось значений функции меньших, чем в $y'_q$, то запускаем локальный метод.

Очищаем множество $V$.

\end{enumerate}

%На рисунке  (\ref{fig:fig2}) отображены линии уровней оптимизационной задачи с отображением текущих  точек испытаний, и на рисунке  (\ref{fig:fig2_2}) отображено соответствующее дерево решений.

Все полученные результаты были добавлены в систему globalizer \cite{fio_bib18}.
Если говорить более конкретно, то была реализована объединенная схема с запуском локальных методов в нужный момент, для чего была внедрена система  построения дерева решений, равномерной сетки и прочих сопутствующих вычислений.  
Ниже приведена блок-схема того, как данная схема была объединена с АГП.

\begin{figure}[ht!]

	\begin{center}
		\begin{minipage}[h]{0.9\linewidth}
			\includegraphics[width=1\linewidth]{figure/fig3.png}
			\caption{Блок-схема объединения АГП и метода Хука-Дживса с использованием дерева решений} %% подпись к рисунку
			\label{fig:fig3}
		\end{minipage}
	\end{center}
\end{figure}

\section{Эксперименты}

Вычислительные эксперименты проводились на персональном компьютере со следующими характеристиками: Операционная система Windows 11,  Процессор: Intel(R) Core™ i5-8250U CPU @ 1.60 GHz, 12 Gb RAM.
В работе \cite{fio_bib13, fio_bib17} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.

В таблице \ref{tab:1} отражено среднее число итераций , которые выполнил метод при решении серии задач из данных классов. Символ «>» отражает ситуацию, когда не все задачи класса были решены каким-либо методом. Количество нерешенных задач указано в скобках.

\begin{table}[!ht]
    \caption{Среднее число итераций}
    \label{tab:1}
    \centering
    \begin{tabular}{|l|l|l|l|l|}
    \hline
        N & Problem class & DIRECT & DIRECTl & АГП  \\ \hline
        4 & Simple & >47282 (4) & 18983 & 11953  \\ \hline
        ~ & Hard & >95708 (7) & 68754 & 25263  \\ \hline
        5 & Simple & >16057 (1) & 16758 & 15920  \\ \hline
        ~ & Hard & >217215 (16) & >269064 (4) & >148342 (4)  \\ \hline
    \end{tabular}
\end{table}

Как видно из таблицы 1, АГП превосходит методы DIRECT и DIRECTl на всех классах задач по среднему числу итераций.

Ниже приведены результаты сравнения двух алгоритмов – стандартного АГП и его объединение с методом Хука-Дживса, при использовании деревьев решений. Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{fio_bib19}. Критерием остановки служило попадание точки испытания в эпсилон окрестность истинного глобального минимума. 
При этом измерялось не только среднее число проводимых испытаний, но и среднее число точек испытаний.

В таблице \ref{tab:2} представлено среднее количество итераций, и среднее число точек испытаний, проводимое двумя методами: алгоритмом глобального поиска (АГП) и объединением АГП и локального метода с использованием деревьев решений (Деревья решений) на задачах разной размерности ($N$) и разного класса (Класс задачи). Значение в скобках (если оно есть) показывает количество не решенных задач. Это означает, что было достигнут максимум по числу проводимых итераций, но в окрестность глобального минимума точка так и не попала.

\begin{table}[h!]
    \caption{Среднее число итераций и среднее число испытаний, проводимое разными алгоритмами}
    \label{tab:2}
    \centering
    \begin{tabular}{|c|c|c|c|c|c|}
    \hline

        N & Класс задачи & \multicolumn{2}{c|}{Среднее число итераций} & \multicolumn{2}{c|}{Среднее число точек испытаний} \\ \hline
          & ~ & АГП & Деревья решений & АГП & Деревья решений \\ \hline
          & Simple & 2350.0 & 247.4 & 2348.0 & 390.0  \\ \hline
        2  & Hard & 4732.3 & 353.6 & 4730.3 & 1101.1  \\ \hline
          & Simple & 2115.5 & 579.5 & 2113.5 & 2441.4  \\ \hline
        3  & Hard & 5347.4 & 588.1 & 5345.4 & 2532.7  \\ \hline
          & Simple & 12168.9 & 768.2 & 12166.9 & 3582.2  \\ \hline
        4  & Hard & 25636.5 & 960.0 & 52634.5 & 5171.6  \\ \hline
          & Simple & 20633.5 & 978.0 & 20631.5 & 4840.4  \\ \hline
        5  & Hard & 161094.9 (4) & 1217.0 & 161092.9 (4) & 7115.5  \\ \hline
    \end{tabular}
\end{table}

Итак, ускорение по числу итераций неоспоримое, причем на задачах любой размерности. Среднее число точек испытаний тоже, практически везде, уменьшилось, но эти значения могут быть сильно выше, чем среднее число испытаний. Это связано с тем, что в точках испытаний учитываются и те точки, которые в процессе своей работы ставил локальный метод, а их может быть довольно много.

\section{Заключение}

Методы глобальной оптимизации являются областью активных научных исследований.  В результате работы удалось успешно объединить алгоритм глобального поиска вместе с поиском локального минимума методом Хука-Дживса. С целью экспериментального подтверждения теоретических свойств рассматриваемого алгоритма были так же проведены вычислительные эксперименты на серии из сотни тестовых задач.

%
% ---- Bibliography ----
%
\renewcommand{\refname}{Список литературы}
\bibliographystyle{IEEEtran}   %%link to spmpsci.bst
\bibliography{article}        %%link to article.bib

\section*{Сведения об авторах}

\paragraph{Лебедев Илья Генадьевич} --- заведующий лабораторией, Нижегородский государственный университет им. Н.И. Лобачевского  e-mail: \url{ilya.lebedev@itmm.unn.ru}. Область профессиональных интересов: методы глобальной и локальной оптимизации, параллельные алгоритмы, программирование для графических процессоров, CUDA. Является автором и соавтором более 30 научных работ в указанных областях.

\paragraph{Д. И. Силенко} --- окончил бакалавриат института информационных технологий, математики и механики Нижегородского государственного университета им. Н. И. Лобачевского в 2021 году. В настоящее время является магистрантом института информационных технологий, математики и механики (направление - прикладная математика и информатика). Область научных интересов включает методы глобальной и локальной оптимизации, а также технологии параллельных вычислений и параллельные алгоритмы. e-mail: \url{dmsi@mail.ru}. Является автором или соавтором нескольких научных работ в указанных областях.

\paragraph{Контактный номер телефона:}{+7(910)898-02-39}

\section*{Authors' bio}

\paragraph{Ilya~G.~Lebedev} --- head of laboratory, Lobachevsky State University of Nizhny Novgorod. e-mail: \url{ilya.lebedev@itmm.unn.ru}. His research interests include algorithms of global and local optimization, parallel algorithms, GPU programming, CUDA. He has authored or coauthored more than 30 papers in these areas.

\paragraph{D. I. Silenko} --- Graduated the Institute of Information Technology, Mathematics and Mechanics of the Lobachevsky State University of Nizhni Novgorod in 2021. He is currently a master student of the Institute of Information Technology, Mathematics and Mechanics. His research interests are algorithms of global and local optimization, parallel algorithms and parallel technologies. e-mail: \url{dmsi@mail.ru}. He has authored or coauthored several papers in these areas.

\end{document}

### 2023-01-22 2023/PAVT_Ufa/PAVT2023.tex
first version

%%%%%%%%%%%%%%%%%%%% author.tex 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{svproc}
%
% RECOMMENDED 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%
\usepackage{graphicx}
\usepackage{marvosym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cite}

\usepackage[russian]{babel}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\usepackage{hyperref}
\def\UrlFont{\rmfamily}

\def\orcidID#1{\unskip$^{[#1]}$}
\def\letter{$^{\textrm{(\Letter)}}$}

\begin{document}
\mainmatter              % start of a contribution
% Решение обратных задач химической кинетки с помощью асинхронного алгоритма глобального поиска
% Решение обратных задач химической кинетки с помощью смешанного локально-глобального поискового алгоритма 
%\title{Solving the Inverse Problems of Chemical Kinetics Using the Asynchronous Global Optimization Algorithm}
\title{Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as Catalyst Using the Asynchronous Global Optimization Algorithm}
\titlerunning{Solving the Inverse Problems of Chemical Kinetics}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{
Irek Gubaydullin$^{1,2}$\and
Leniza Enikeeva$^{2,3}$\orcidID{0000-0003-4219-4870}
\and
Konstantin Barkalov$^4$ \letter \orcidID{0000-0001-5273-2471}
\and
Ilya Lebedev$^4$\orcidID{0000-0002-8736-0652} 
}

%
\authorrunning{I. Gubaydullin et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Konstantin Barkalov and Ilya Lebedev }
%

\institute{$^1$Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia\\$^2$Ufa State Petroleum Technological University, Ufa, Russia \\$^3$Novosibirsk State University, Novosibirsk, Russia\\$^4$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
\email{leniza.enikeeva@yandex.ru},
\email{konstantin.barkalov@itmm.unn.ru},
\email{ilya.lebedev@itmm.unn.ru}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}

The paper considers the application of parallel computing technology to the simulation of catalytic chemical reaction, which is widely used in the modern automobile industry to produce gasoline with high octane number. As a chemical reaction, the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid is assumed. To simulate a chemical process, it is necessary to develop a kinetic model of the process, that is, to determine the kinetic parameters. To do this, the inverse problem of chemical kinetics is solved, which predicts the values of kinetic parameters based on laboratory data. From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem. A parallel information-statistical global search algorithm was used to solve it. The use of the parallel algorithm has significantly reduced the search time to find the optimum. The found optimal parameters of the model made it possible to adequately simulate the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid.

%обновить ключевые слова
\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Chemical kinetics $\cdot$ Inverse problems }
\end{abstract}

\section{Introduction}

%Часть УГНТУ
Currently, there is a tendency to improve the environmental characteristics of automobile fuel while maintaining a high octane number. Sulfuric acid alkylation of isobutane with olefins makes it possible to obtain a high-octane component of gasoline with a minimum content of aromatic hydrocarbons. The alkylate, which is produced by the alkylation of isobutane with C3-C5 olefins in the presence of strong acid, has the advantages of high octane number, low vapor pressure, and zero content of olefins and aromatics that allow it to be a desirable blending component for high-quality gasoline. Alkylates will continue to act as a desirable blending component for high-quality gasoline as the quality of gasoline continues to increase [ссылка на китайскую статью]. Therefore, it is a significant process of the modern refinery [ссылка на catalyst]. To optimize the chemical process in industry, it is necessary to first develop its model, which in this case means building a mathematical model of the chemical process and then its kinetic model, that is, numerically calculate the kinetic constants of the reaction. 
%Стыковочный абзац
Как правило, найти кинетические константы реакций аналитическим путем не представляется возможным, поэтому возникает необходимость в разработке и использовании численных методов их поиска (см., например, [ссылки (Наша совместная статья Суперкомп дни 2021 ??)]). В этом случае критерий качества найденного решения (целевая функция) не имеет явного аналитического описания, но допускает алгоритмическое представление и требует значительных затрат на расчет. Кроме того, в обратных задачах химической кинетики целевая функция может быть существенно многоэкстремальной, т.е. иметь много локальных экстремумов наряду с глобальным. 
%Часть ННГУ
Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) существенно отличаются от методов локального поиска (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}). Как правило, методы локальной оптимизации не могут покинуть области притяжения локального экстремума и не находят глобальный оптимум. А использование параметров модели, соответствующих найденному локальному решению, может оказаться недостаточным, поскольку глобальное решение может дать существенный выигрыш по сравнению с локальным.

Разнообразие возникающих задач глобальной оптимизации влечет за собой разнообразие подходов к их решению.
Методы решения задач глобальной оптмизации можно условно разделить на два класса: метаэвристические и детерминированные. Метаэвристические алгоритмы, как правило, основаны на имитации процессов, протекающих в живой природе.
Примерами метаэвристических алгоритмов являются simulated annealing, evolution and genetic algorithms и т.д. (see, e.g., \cite{Battiti2009,Eiben2015}). В силу своей относительной простоты метаэвристические алгоритмы более популярны у исследователей, чем детерминированные.  Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может распологаться далеко от глобального \cite{Kvasov2018}.

Возможность построения детерминированных методов глобального поиска, отличающихся от перебора по сетке и метаэвристических алгоритмов, связана с наличием и учетом  априорных предположений о свойствах функций задачи. Такие предположения играют ключевую роль при разработке эффективных алгоритмов глобальной оптимизации и служат основным математическим инструментом для получения оценок глобального решения.

Одним из естественных допущений о задаче является предположение об ограниченности относительных изменений значений целевой функции. Такое допущение связано с тем, что отношения приращений функции к соответствующим приращениям ее аргумента обычно не могут превышать некоторого порога, определяемого ограниченной энергией изменений в моделируемой системе. В таком случае, как известно, функции называются липшицевыми, а сама задача -- задачей липшицевой глобальной оптимизации.

В данной статье отражены результаты применения параллельных методов липшицевой оптимизации, разрабатываемых в ННГУ им. Н.И. Лобачевского, к решению обратных задач химической кинетики. 
Основная часть статьи имеет следующую структуру. Описание математической модели исследуемой химической реакции приведено в Section 2. Формальная постановка задачи липшицевой глобальной оптимизации приводятся в Section 3. В Section 4 кратко изложена схема разрабатываемого авторами асинхронного параллельного алгоритма для решения задач указанного класса. Результаты численного решения обратной задачи химической кинетики обсуждаются в Section 5.

\section{Problem Statement}\label{Sec_math_mod}
%Содержательная постановка задачи

Рассмотрим математическую модель реакции алкилирования изобутана олефинами в присутствии серной кислоты, которая представляет собой систему обыкновенных нелинейлых дифференциальных уравнений (\ref{eq:one}) -- (\ref{eq:twelve}).

\begin{equation}
  \frac{dc_1}{dt} = -k_1c_1 + k_2c_3 - k_3c_1c_3 - k_7c_1c_2c_4 - k_{11}c_1 + k_{14}c_{11}
  \label{eq:one}
\end{equation}

\begin{equation}
  \frac{dc_2}{dt} = -k_4c_2c_4 - k_6c_2c_5 - k_7c_1c_2c_4 - k_{15}c_{11}c_2c_4
  \label{eq:two}
\end{equation}

\begin{equation}
  \frac{dc_3}{dt} = k_1c_1 + k_4c_2c_5 - k_3(c_1 + c_{11})c_3 - k_{5}c_{12}c_3 - k_2c_3 + k_7c_1c_2c_4 + k_{15}c_{11}c_2c_4
  \label{eq:three}
\end{equation}

\begin{equation}
  \frac{dc_4}{dt} = k_3(c_1 + k_{11})c_3 - k_4c_2c_4 - k_{7}c_{1}c_2c_4 - k_{15}c_{11}c_2c_4
  \label{eq:four}
\end{equation}

\begin{equation}
  \frac{dc_5}{dt} = k_5c_{12}c_3 - k_{6}c_2c_5
  \label{eq:five}
\end{equation}

\begin{equation}
  \frac{dc_6}{dt} = k_4c_{2}c_4
  \label{eq:six}
\end{equation}

\begin{equation}
  \frac{dc_7}{dt} = k_6c_{2}c_5 - k_{10}c_7
  \label{eq:seven}
\end{equation}

\begin{equation}
  \frac{dc_8}{dt} = k_7c_{1}c_2c_4 + k_{15}c_{11}c_2c_4 + k_9c_9c_{10} - k_8c_8
  \label{eq:eight}
\end{equation}

\begin{equation}
  \frac{dc_9}{dt} = k_8c_{8} - k_{9}c_{9}c_{10}
  \label{eq:nine}
\end{equation}

\begin{equation}
  \frac{dc_{10}}{dt} = k_8c_{8} - k_{9}c_{9}c_{10}
  \label{eq:ten}
\end{equation}

\begin{equation}
  \frac{dc_{11}}{dt} = -k_3c_{11}c_3 - k_{15}c_{11}c_{2}c_4 + k_{11}c_1 + k_{12}c_{12} - k_{13}c_{11} - k_{14}c_{11}
  \label{eq:eleven}
\end{equation}

\begin{equation}
  \frac{dc_{12}}{dt} = -k_5c_{12}c_3 + k_{13}c_{11} - k_{12}c_{12}
  \label{eq:twelve}
\end{equation}

The initial conditions are $t = 0, c_1 = c_1^0; c_2=c_2^0; c_3 = 0; c_4 = 0; c_5= 0; c_6 = 0; c_7 = 0; c_8 = 0; c_9 = 0; c_{10} = 0; c_{11}=c_{11}^0; c_{12} = c_{12}^0$.
The corresponding species in eqs (\ref{eq:one}) -- (\ref{eq:twelve}) are 1, iC4H8; 2, iC4, 3, iC4+; 4, TMPs+; 5, DMHs+; 6, TMPs; 7, DMHs; 8, HEs; 9, iCx+; 10, iCy=; 11, 2-C4H8; 12, 1-C4H8.

В качестве экспериментальных данных взяты данные со статьи [], которые представляют собой изменение концентраций компонентов реакции во времени при различных температурах; в качестве примера экспериментальных данных для сырья \#1 при температуре 276.2 K представлены данные в таблице \ref{table1}.

\begin{table}
\label{table1}
\caption{Experimental Data}
\begin{center}
\begin{tabular}{ccccccc}
\hline
%\multicolumn{1}{l}{\rule{0pt}{12pt}
%                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
& 1 min & 2 min & 5 min & 10 min & 15 min & 20 min  \\
\hline\rule{0pt}{12pt}
DMH & 0.12 & 0.11 & 0.1	& 0.1 &	0.095 &	0.09  \\
TMP & 0.54 & 0.65 & 0.69 & 0.69 & 0.7 & 0.705 \\[2pt]
\hline
\end{tabular}
\end{center}
\end{table}

Таким образом, решая систему (\ref{eq:one}) -- (\ref{eq:twelve}) с соответсвующими начальными данными, мы получим изменение рассчитанных концентраций компонентов реакции по времени. ...
Необходимо минимизировать следующую целевую фукнцию:

\begin{equation}
  F = \sum ...
  \label{eq:objective_func}
\end{equation}

\section{Parallel Algorithm for Solving Global Optimization Problems }\label{Sec_GSA}

\subsection{Global Optimization Problem}

Как уже говорилось выше, задача идентификации значений параметров математической модели может рассматриваться как Lipschitz global optimization problem. 
С формальной точки зрения  данная задача является задачей математического программирования вида
\begin{gather}
 \varphi^* = \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problemN}\\
 D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\} \label{D},
\end{gather}
где $a,\; b$ есть заданные векторы, $a,b\in R^N$, $\varphi(y)$ целевая функция, которая удовлетворяет the Lipschitz condition
\begin{equation}\label{Lip}
\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D.
\end{equation}

% До схемы характеристического алгоритма - переписать

Функция $\varphi(y)$ предполагается многоэкстремальной и заданной в виде ``black box'' (т.е. в виде некоторой вычислительной процедуры, на вход которой подается вектор параметров, а на выходе наблюдается соответствующее значение функции). Кроме того, подразумевается, что каждое trial (т.е. вычисление значение функции в точке области поиска) является трудоемкой операцией и требует значительных вычислительных ресурсов. Как отмечалось во введении, такая постановка задачи полностью соответствует the inverse problem of chemical kinetics.

The Lipschitz condition (\ref{Lip}) может быть использовано для оценки глобального минимума функции на интервале, а знание константы Липшица позволяет конструировать алгоритмы глобального поиска и доказывать условия их сходимости (see, for example, \cite{Strongin2000}).

Одной из основных трудностей в решении многомерных задач глобальной оптимизации является рост вычислительных затрат при увеличении размерности задачи. Уменьшение числа испытаний функции при сохранении точности решения возможно за счет более полного использования априорных предположений о целевой функции, что приводит к адаптивным последовательным алгоритмам оптимизации.

Такими методами являются, например, non-uniform space covering method \cite{Evtushenko2013} and simplicial partitions method \cite{Zilinskas2010}. Данные подходы были успешно использованы и для построения параллельных оптимизационных методов \cite{Evtushenko2009,Paulavicius2011}. Другим адаптивным подходом к решению многомерной задачи (\ref{problemN}) является ее редукция к одной или нескольким одномерным задачам с последующим применением одномерных алгоритмов. Такая редукция может быть осуществлена, например, при помощи nested optimization scheme \cite{Grishagin2018} или при помощи Peano-Hilbert curves \cite{Barkalov2018}. Последний из указанных подходов и используется в настоящей работе.

Используя непрерывное однозначное отображение (Peano-Hilbert curve) $y(x)$ отрезка $[0,1]$ вещественной оси на гиперинтервал $D$ из (\ref{D}) можно свести многомерную задачу (\ref{problemN}) к одномерной задаче
\[
\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}},
\]
where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
\[
\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}
\]
where $ H=2 L \sqrt{N+3}$.
Вопросы численного построения различных аппроксимаций Peano-Hilbert curve рассмотрены в \cite{Strongin2000,Sergeyev2013}.

Таким образом, поисковое испытание в некотрой точке $x'\in[0,1]$ будет включать в себя сначала the construction of the image $y'=y(x')$, и лишь затем вычисление значения функции $ z' = \varphi(y')$.

%Характеристический алгоритм

A numerical optimization method belongs to the class of characteristical algorithms if the algorithmic scheme of this method can be described as follows.

At the first iteration the trials are performed at the points $x^1 = 0$ and $x^2 = 1$. Then any next $(k+1)$, $k \geq 2$, iteration og global search is executed in the following way.

\begin{enumerate}
	\item 
Renumber points of previous trials by increasing their $x$-coordinates (the new order is marked with subscript)
\[
0=x_0<x_2<...<x_{k}=1.
\]

	\item 
Calculate for every interval $(x_{i-1},x_i), 1\leq i\leq k$, a value $R(i)$ called \textit{the interval characteristic}. In general case, $R(i)$ can depend on the points $x^i$ and the trial rezults $z^i=f(x^i), 1 \leq i \leq k$.

	\item 
Find the interval $(x_{t-1},x_t)$ which has the largest characteristic $R(t)$, i.e.
\[
R(t) = \max \left\{ R(i): \; 1\leq i\leq k \right\}.
\]

	\item 
Examine the stop condition
\[
\Delta_t \leq \epsilon ,
\]
where $\epsilon>0$ is a given accuracy. If the stop condition is satisfied, then the global search has to be terminated; otherwise the calculations should continue.

	\item 
Select a point $x^{k+1}$ of the current iteration within the interval $(x_{t-1},x_t)$ in accordance with some rule $S(t)$, i.e.
\[
x^{k+1} = S(t)\in(x_{t-1},x_t).
\]

	\item 
Calculate the function value $z^{k+1} = f(x^{k+1})$

	\item 
Evaluate a global minimum estimate

	\item 
Increase the iteration number $k=k+1$ and proceed to the next iteration.
\end{enumerate}

As a possible interpretation of this scheme, we can consider an interval characteristic $R(i)$ as a measure of the global minimum being within the interval
$(x_{i-1},x_i)$.
To construct a concrete form of the interval characteristic we can use a lower envelope (or minorant) of the function to be minimized or a mathematical expectation of function values, etc. 
Most well-known global optimization methods can be formulated in accordance with this characteristical scheme, e.g.,
\begin{itemize}
\item a uniform grid method with a successively reduced step;
\item a random search (Monte-Carlo) algorithm;
\item the Piyavskii method;
\item one-step bayesian methods proposed by Kushner and Zilinskas;
\item information algorithms proposed by Strongin.
\end{itemize}
All these methods are based on different mathematical models but are presented in the general characteristical scheme.

It should be also noted that the length of the interval with the largest characteristic is examined at the stop condition.
It is possible if the optimization method converges to the global minimum.

Details of these methods are given in the sources cited. %Здесь же отметим основные моменты

For the Piyavskii method the interval characteristic R(i) is an estimate (with the inverse sign) of the least value of the objective function $f(x)$ within the
interval $(x_{i-1},x_i)$. As a result, the point of a new trial is taken within the interval containing the estimate of the least value of $f(x)$ over the search domain.

The Kushner technique and the Zilinskas method have been constructed in the framework of the approach when the objective function is considered as
a sample of some Wiener process. For the Kushner technique the point $x^{k+1}$ of the current iteration is the most probable point at which the function value
$f(x^{k+1})$ is not greater than the value
\[
z_k^* -\gamma (z_k^+-z_k^*),
\]
where $z_k^+$ and $z_k^*$ are estimates of maximum and minimum function values respectively. For the Zilinskas method $x^{k+1}$ is the point where the maximum average improvement of the current estimate of the global extremum is expected.

The Strongin algorithm has been constructed in the framework of the information approach to global optimization (see Strongin (1978)). This method has an adaptive scheme to evaluate a numerical estimate of the unknown Lipschitz constant.

%Связка про распараллеливание

\subsection{Parallel Asynchronous Global Search Algorithm}

В предлагаемом подходе схема распараллеливания соответствует по принципу ``master/worker''. В процессе-мастере выполняется алгоритм глобального поиска, в котором проводится накопление поисковой информации, оценка на ее основе константы Липшица для целевой функции, определение точек новых испытаний и распределение их по процессам-рабочим. Процессы-рабочие получают от мастера точки, проводят в них новые испытания и отсылают мастеру результаты испытаний.

Будем считать, что на каждой итерации процесс-мастер вычисляет одну точку нового испытания и передает ее для проведения испытания в процесс-рабочий. 
Так как проведение испытания рабочим является значительно более трудоемкой операцией, чем выбор новой точки мастером, то это исключает простой процессов-рабочих. А процесс-мастер будет загружен обработкой поисковой информации, что минимизирует простой мастера.
Тем самым асинхронная схема будет обеспечивать хорошую эффективность распараллеливания.
При этом общее число выполненных испытаний каждым процессом-рабочим будет зависеть от трудоемкости проведения конкретного испытания и не может быть оценено заранее.

При описании параллельного алгоритма будем предполагать, что в нашем распоряжении имеется $p+1$ вычислительный процесс: один мастер и $p$ рабочих.

В начале поиска процесс-мастер (будем считать, что это процесс No 0) инициирует параллельное проведение $p$ испытаний в $p$ различных точках области поиска, две из которых являются граничными, а остальные -- внутренними, т.е. в точках $\{y(x^1), y(x^2), ...,y(x^p)\}$, где $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.

Предположим теперь, что выполнено $k$ испытаний (в частности, $k$ может быть равно 0), и процессы-рабочие проводят испытания в точках $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.

Каждый процесс-рабочий, завершивший свое испытание в некоторой точке (не ограничивая общности, будем считать, что это точка $y(x^{k+1})$, соответствующая процессу No 1), пересылает процессу-мастеру результаты испытания. 
Отметим, что в данном случае мы будем иметь set of preimages of the trial points
\[
I_k = \left\{ x^{k+1},x^{k+2},...,x^{k+p} \right\},
\]
в которых испытания уже начались, но еще не завершены.

Получив от рабочего процесса результаты испытания в точке $y(x^{k+1})$ мастер, в свою очередь, выбирает для него точку нового испытания $x^{k+p+1}$ в соответствии с правилами, описанными ниже.

Step 1. Перенумеровать в порядке возрастания (нижним идексом) set of preimages of the trial points 
\[
X_k = \left\{x^1, x^2,...,x^{k+p} \right\},
\]
которое содержит все preimages, в которых либо проведены, либо проводятся испытания, т.е.
\[
0=x_1<x_2<...<x_{x+p}=1.
\]
Step 2. Вычислить значения 
\[
M_1=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin I_k, x_i \notin I_k, 2\leq i\leq k+p \right\},
\]
\[
M_2=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in I_k, 2\leq i < k+p \right\},
\]
\[
M=\max\{M_1,M_2\},
\]
где $z_i=\varphi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$. Значения $z_i$ в точках $x_i \in I_k$ являются неопределенными, т.к. испытания в точках $x_i \in I_k$ еще не завершены. Если значение $M$ получилось равным 0, то положить $M=1$.

Step 3. Каждому интервалу $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, поставить в соответствие число $R(i)$, которое называется характеристикой интервала и вычисляется по формуле
\[
R(i)=rM\Delta_i+\frac{(z_i-z_{i-1})^2}{rM\Delta_i}-2(z_i+z_{i-1}),
\]
где $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.

Step 4. Выбрать интервал $[x_{t-1},x_t]$, которому соответствует максимальная характеристика, т.е.
\[
R(t) = \max \left\{ R(i): \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p \right\}.
\]

Step 5. Определить прообраз $x^{k+p+1} \in (x_{t-1},x_t)$ точки нового испытания в соответствии с формулой
\[
x^{k+p+1} = \frac{x_{t}+x_{t-1}}{2} - \mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_{t}-z_{t-1}\right|}{M}\right]^N.
\]

Сразу после вычисления точки очередного испытания $y^{k+p+1} = y(x^{k+p+1})$ процесс-мастер добавляет ее в множество $I_k$ и пересылает процессу-рабочему, который инициирует проведение испытания в ней.

Процесс-мастер завершает работу алгоритма, если выполняется одно из двух условий: $\Delta_{t}<\epsilon$, $k+p>K_{max}$.
Вещественное число $\epsilon>0$ и целое число $K_{max}>0$ являются параметрами алгоритма и соответствуют точности поиска решения и максимальному числу испытаний.

Описанный параллельный асинхнонный алгоритм основан на последовательном information global search algorithm. Теоретическое обоснование сходимости алгоритма приведено в \cite{Strongin2000}. Здесь же приведены схемы синхронного распараллеливания, использовавшиеся ранее при решении ряда прикладных задач \cite{Kalyulin2017,Modorskii2016}.
Новизна данной работы заключается в практической реализации и применении асинхронной схемы распалаллеливания, обладающей большей эффективностью при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска. Это подтверждается результатами экспериментов, описанными в следующем разделе.

\section{Numerical Experiments}\label{Sec_Exp}

\section{Conclusions and Future Work}

\medskip

\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204 and by RFBR, project No.\,19-37-60014.

%
% ---- Bibliography ----
%
\bibliographystyle{spmpsci}
\bibliography{bibliography}{}

\end{document}

### 2023-01-23 2023/PAVT_Ufa/PAVT2023.tex
first part in Russian

@@
 Разнообразие возникающих задач глобальной оптимизации влечет за собой разнообразие подходов к их решению.
-Методы решения задач глобальной оптмизации можно условно разделить на два класса: метаэвристические и детерминированные.
+Методы решения задач глобальной оптимизации можно условно разделить на два класса: метаэвристические и детерминированные.
 Метаэвристические алгоритмы, как правило, основаны на имитации процессов, протекающих в живой природе.
@@
 В силу своей относительной простоты метаэвристические алгоритмы более популярны у исследователей, чем детерминированные.
-Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может распологаться далеко от глобального \cite{Kvasov2018}.
+Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может располагаться далеко от глобального \cite{Kvasov2018}.
 
-Возможность построения детерминированных методов глобального поиска, отличающихся от перебора по сетке и метаэвристических алгоритмов, связана с наличием и учетом  априорных предположений о свойствах функций задачи.
-Такие предположения играют ключевую роль при разработке эффективных алгоритмов глобальной оптимизации и служат основным математическим инструментом для получения оценок глобального решения.
-
-Одним из естественных допущений о задаче является предположение об ограниченности относительных изменений значений целевой функции.
-Такое допущение связано с тем, что отношения приращений функции к соответствующим приращениям ее аргумента обычно не могут превышать некоторого порога, определяемого ограниченной энергией изменений в моделируемой системе.
+Предполагая наличие некоторых дополнительных свойств целевой функции, можно построить эффективные детерминированные методы поиска глобального решения.
+Например, можно предположить, что отношение приращения функции к соответствующему приращению аргумента не может превышать некоторого порога.
 В таком случае, как известно, функции называются липшицевыми, а сама задача -- задачей липшицевой глобальной оптимизации.
 
-В данной статье отражены результаты применения параллельных методов липшицевой оптимизации, разрабатываемых в ННГУ им. Н.И. Лобачевского, к решению обратных задач химической кинетики. 
+Данная статья продолжает цикл работ, в которых проводится исследование и модификация параллельных методов липшицевой оптимизации при их использовании для решения обратных задач химической кинетики. 
 Основная часть статьи имеет следующую структуру.
@@
 Функция $\varphi(y)$ предполагается многоэкстремальной и заданной в виде ``black box'' (т.е.
-в виде некоторой вычислительной процедуры, на вход которой подается вектор параметров, а на выходе наблюдается соответствующее значение функции).
-Кроме того, подразумевается, что каждое trial (т.е.
-вычисление значение функции в точке области поиска) является трудоемкой операцией и требует значительных вычислительных ресурсов.
-Как отмечалось во введении, такая постановка задачи полностью соответствует the inverse problem of chemical kinetics.
+в виде некоторой subroutine, на вход которой подается вектор параметров, а на выходом является вычисленное значение функции).
+Кроме того, подразумевается, что каждое применение этой процедуры (называемое в дальнейшем search trial) является трудоемкой операцией.
+Такая постановка задачи полностью соответствует the inverse problem of chemical kinetics.
 
-The Lipschitz condition (\ref{Lip}) может быть использовано для оценки глобального минимума функции на интервале, а знание константы Липшица позволяет конструировать алгоритмы глобального поиска и доказывать условия их сходимости (see, for example, \cite{Strongin2000}).
+%The Lipschitz condition (\ref{Lip}) может быть использовано для оценки глобального минимума функции на интервале, а знание константы Липшица позволяет конструировать алгоритмы глобального поиска и доказывать условия их сходимости (see, for example, \cite{Strongin2000}).
 
-Одной из основных трудностей в решении многомерных задач глобальной оптимизации является рост вычислительных затрат при увеличении размерности задачи.
+%Одной из основных трудностей в решении многомерных задач глобальной оптимизации является рост вычислительных затрат при увеличении размерности задачи.
 Уменьшение числа испытаний функции при сохранении точности решения возможно за счет более полного использования априорных предположений о целевой функции, что приводит к адаптивным последовательным алгоритмам оптимизации.
 
-Такими методами являются, например, non-uniform space covering method \cite{Evtushenko2013} and simplicial partitions method \cite{Zilinskas2010}.
-Данные подходы были успешно использованы и для построения параллельных оптимизационных методов \cite{Evtushenko2009,Paulavicius2011}.
-Другим адаптивным подходом к решению многомерной задачи (\ref{problemN}) является ее редукция к одной или нескольким одномерным задачам с последующим применением одномерных алгоритмов.
-Такая редукция может быть осуществлена, например, при помощи nested optimization scheme \cite{Grishagin2018} или при помощи Peano-Hilbert curves \cite{Barkalov2018}.
-Последний из указанных подходов и используется в настоящей работе.
+Известен ряд алгоритмов, применимых для решения задач липшицевой оптимизации.
+Среди них можно отметить non-uniform space covering method \cite{Evtushenko2013,Evtushenko2009}, diagonal and simplicial partitions methods \cite{Zilinskas2010,Paulavicius2011}.
+В данной работе мы использовали global search algorithm, proposed by Strongin \cite{Strongin2000}.
+В рамках данного подхода исходная многомерная задача (\ref{problemN}) редуцируется к задаче одномерной оптимизации при помощи Peano-Hilbert curves.
 
-Используя непрерывное однозначное отображение (Peano-Hilbert curve) $y(x)$ отрезка $[0,1]$ вещественной оси на гиперинтервал $D$ из (\ref{D}) можно свести многомерную задачу (\ref{problemN}) к одномерной задаче
+В самом деле, используя непрерывное однозначное отображение (Peano-Hilbert curve) $y(x)$ отрезка $[0,1]$ вещественной оси на гиперинтервал $D$ из (\ref{D}) можно свести многомерную задачу (\ref{problemN}) к одномерной задаче
 \[
@@
 \]
-where $ H=2 L \sqrt{N+3}$.
+with $ H=2 L \sqrt{N+3}$.
 Вопросы численного построения различных аппроксимаций Peano-Hilbert curve рассмотрены в \cite{Strongin2000,Sergeyev2013}.
 
-Таким образом, поисковое испытание в некотрой точке $x'\in[0,1]$ будет включать в себя сначала the construction of the image $y'=y(x')$, и лишь затем вычисление значения функции $ z' = \varphi(y')$.
+Таким образом, search trial в некоторой точке $x'\in[0,1]$ будет включать в себя сначала the construction of the image $y'=y(x')$, и лишь затем вычисление значения функции $ z' = \varphi(y')$.
 
 %Характеристический алгоритм
-
-A numerical optimization method belongs to the class of characteristical algorithms if the algorithmic scheme of this method can be described as follows.
+Примененный нами алгоритм глобального поиска принадлежит классу characteristic algorithms, что существенно упрощает процесс его распараллеливания.
+Напомним, что а numerical optimization method belongs to the class of characteristical algorithms if the algorithmic scheme of this method can be described as follows.
 
@@
 
-В предлагаемом подходе схема распараллеливания соответствует по принципу ``master/worker''.
-В процессе-мастере выполняется алгоритм глобального поиска, в котором проводится накопление поисковой информации, оценка на ее основе константы Липшица для целевой функции, определение точек новых испытаний и распределение их по процессам-рабочим.
-Процессы-рабочие получают от мастера точки, проводят в них новые испытания и отсылают мастеру результаты испытаний.
+Принадлежность GSA к классу caracteristic algorithms существенно упрощает схему его распараллеливания. 
+Как уже было отмечено, характеристика интервала $R(i)$ может рассматриваться как некоторая мера нахождения в нем интервалах точки глобального минимума. 
+Тогда вместо одного лучшего интервала можно выбрать сразу несколько интервалов с наибольшими характеристиками и провести в них очередные испытания параллельно.
+Более того, схема характеристических алгоритмов допускает и асинхронное распараллеливание, которое минимизирует простой процессоров в случае, если трудоемкость проведения испытаний зависит от конкретной точки точках области поиска.
 
-Будем считать, что на каждой итерации процесс-мастер вычисляет одну точку нового испытания и передает ее для проведения испытания в процесс-рабочий. 
-Так как проведение испытания рабочим является значительно более трудоемкой операцией, чем выбор новой точки мастером, то это исключает простой процессов-рабочих.
-А процесс-мастер будет загружен обработкой поисковой информации, что минимизирует простой мастера.
-Тем самым асинхронная схема будет обеспечивать хорошую эффективность распараллеливания.
-При этом общее число выполненных испытаний каждым процессом-рабочим будет зависеть от трудоемкости проведения конкретного испытания и не может быть оценено заранее.
+Рассмотрим более детально асинхронный алгоритм глобального поиска.
+В нем реализована схема распараллеливания вида ``master/worker''.
+В процессе-мастере проводится накопление поисковой информации, оценка на ее основе константы Липшица для целевой функции, определение точек новых испытаний и распределение их по процессам-рабочим.
+Процессы-рабочие получают от мастера точки, проводят в них новые испытания и отсылают мастеру их результаты.
 
@@
 
-В начале поиска процесс-мастер (будем считать, что это процесс No 0) инициирует параллельное проведение $p$ испытаний в $p$ различных точках области поиска, две из которых являются граничными, а остальные -- внутренними, т.е.
+В начале поиска процесс-мастер (пусть это будет процесс No 0) инициирует параллельное проведение $p$ испытаний в $p$ различных точках области поиска, две из которых являются граничными, а остальные -- внутренними, т.е.
 в точках $\{y(x^1), y(x^2), ...,y(x^p)\}$, где $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.
@@
 
-Каждый процесс-рабочий, завершивший свое испытание в некоторой точке (не ограничивая общности, будем считать, что это точка $y(x^{k+1})$, соответствующая процессу No 1), пересылает процессу-мастеру результаты испытания.
+Если процесс-рабочий завершил проведение испытания в некоторой точке (пусть это будет точка $y(x^{k+1})$, соответствующая процессу No 1), то он пересылает процессу-мастеру результаты испытания. 
 Отметим, что в данном случае мы будем иметь set of preimages of the trial points
@@
 
-Получив от рабочего процесса результаты испытания в точке $y(x^{k+1})$ мастер, в свою очередь, выбирает для него точку нового испытания $x^{k+p+1}$ в соответствии с правилами, описанными ниже.
+Получив от рабочего процесса результаты испытания в точке $y(x^{k+1})$ мастер, в свою очередь, выбирает для него точку нового испытания $x^{k+p+1}$ в соответствии с правилами, которые соответствуют схеме характеристического алгоритма.
 
-Step 1. Перенумеровать в порядке возрастания (нижним идексом) set of preimages of the trial points 
+1. Перенумеровать в порядке возрастания (нижним индексом) set of preimages of the trial points 
 \[
@@
 \]
-Step 2. Вычислить значения 
+2. Вычислить значения 
 \[
@@
 
-Step 3.
+3.
 Каждому интервалу $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, поставить в соответствие число $R(i)$, которое называется характеристикой интервала и вычисляется по формуле
@@
 
-Step 4. Выбрать интервал $[x_{t-1},x_t]$, которому соответствует максимальная характеристика, т.е.
+4. Выбрать интервал $[x_{t-1},x_t]$, которому соответствует максимальная характеристика, т.е.
 \[
@@
 
-Step 5. Определить прообраз $x^{k+p+1} \in (x_{t-1},x_t)$ точки нового испытания в соответствии с формулой
+5. Определить прообраз $x^{k+p+1} \in (x_{t-1},x_t)$ точки нового испытания в соответствии с формулой
 \[
@@
 
-Процесс-мастер завершает работу алгоритма, если выполняется одно из двух условий: $\Delta_{t}<\epsilon$, $k+p>K_{max}$.
-Вещественное число $\epsilon>0$ и целое число $K_{max}>0$ являются параметрами алгоритма и соответствуют точности поиска решения и максимальному числу испытаний.
+Процесс-мастер завершает работу алгоритма при выполнении одного из двух условий: $\Delta_{t}<\epsilon$, $k+p>K_{max}$.
+Первое из них соответствует остановке алгоритма по точности, второе --- по числу испытаний. 
+Вещественное число $0<\epsilon<1$ и целое число $K_{max}>0$ являются параметрами алгоритма.
 
-Описанный параллельный асинхнонный алгоритм основан на последовательном information global search algorithm.
+%Описанный параллельный асинхнонный алгоритм основан на последовательном information global search algorithm.
 Теоретическое обоснование сходимости алгоритма приведено в \cite{Strongin2000}.
 Здесь же приведены схемы синхронного распараллеливания, использовавшиеся ранее при решении ряда прикладных задач \cite{Kalyulin2017,Modorskii2016}.
-Новизна данной работы заключается в практической реализации и применении асинхронной схемы распалаллеливания, обладающей большей эффективностью при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска.
+%Новизна данной работы заключается в практической реализации и применении асинхронной схемы распалаллеливания, обладающей большей эффективностью при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска.
 Это подтверждается результатами экспериментов, описанными в следующем разделе.

### 2023-01-24 2023/PAVT_Ufa/PAVT2023.tex
first part in Russian

@@
 \and
-Ilya Lebedev$^4$\orcidID{0000-0002-8736-0652} 
+Marina Usova$^4$\orcidID{XXX} 
 }
@@
 %Стыковочный абзац
-Как правило, найти кинетические константы реакций аналитическим путем не представляется возможным, поэтому возникает необходимость в разработке и использовании численных методов их поиска (см., например, [ссылки (Наша совместная статья Суперкомп дни 2021 ??)]).
-В этом случае критерий качества найденного решения (целевая функция) не имеет явного аналитического описания, но допускает алгоритмическое представление и требует значительных затрат на расчет.
-Кроме того, в обратных задачах химической кинетики целевая функция может быть существенно многоэкстремальной, т.е.
-иметь много локальных экстремумов наряду с глобальным.
+
+Так как математической моделью реакции является система дифференциальных уравнений, найти найти значения констант, входящих в него, можно только численно (см., например, \cite{RSD}).
+Отметим, что в таких задачах целевая функция, как правило, является многоэкстремальной, т.е.
+имеет много локальных экстремумов наряду с глобальным.
 %Часть ННГУ
 Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) существенно отличаются от методов локального поиска (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}).
-Как правило, методы локальной оптимизации не могут покинуть области притяжения локального экстремума и не находят глобальный оптимум.
-А использование параметров модели, соответствующих найденному локальному решению, может оказаться недостаточным, поскольку глобальное решение может дать существенный выигрыш по сравнению с локальным.
-
-Разнообразие возникающих задач глобальной оптимизации влечет за собой разнообразие подходов к их решению.
-Методы решения задач глобальной оптимизации можно условно разделить на два класса: метаэвристические и детерминированные.
-Метаэвристические алгоритмы, как правило, основаны на имитации процессов, протекающих в живой природе.
-Примерами метаэвристических алгоритмов являются simulated annealing, evolution and genetic algorithms и т.д.
+С алгоритмической точки зрения их можно разделить на два класса: метаэвристические и детерминированные.
+Метаэвристические алгоритмы, в основном, основаны на имитации процессов, протекающих в живой природе.
+Характерными примерами таких алгоритмов являются simulated annealing, evolution and genetic algorithms и т.д.
 (see, e.g., \cite{Battiti2009,Eiben2015}).
@@
 Например, можно предположить, что отношение приращения функции к соответствующему приращению аргумента не может превышать некоторого порога.
-В таком случае, как известно, функции называются липшицевыми, а сама задача -- задачей липшицевой глобальной оптимизации.
+В таком случае функции называются липшицевыми, а сама задача -- задачей липшицевой глобальной оптимизации .
+Данная статья продолжает цикл работ, в которых проводится исследование и модификация параллельных методов липшицевой оптимизации, предложенных в \cite{Strongin2000}, при их использовании для решения обратных задач химической кинетики.
 
-Данная статья продолжает цикл работ, в которых проводится исследование и модификация параллельных методов липшицевой оптимизации при их использовании для решения обратных задач химической кинетики. 
 Основная часть статьи имеет следующую структуру.
 Описание математической модели исследуемой химической реакции приведено в Section 2.
-Формальная постановка задачи липшицевой глобальной оптимизации приводятся в Section 3.
-В Section 4 кратко изложена схема разрабатываемого авторами асинхронного параллельного алгоритма для решения задач указанного класса.
-Результаты численного решения обратной задачи химической кинетики обсуждаются в Section 5.
+Формальная постановка задачи липшицевой глобальной оптимизации и общая схема поисковых алгоритмов приводятся в Section 3.
+В Section 4 изложена схема предлагаемого асинхронного параллельного алгоритма для решения многоэкстремальных задач.
+Результаты численного решения the inverse problem of chemical kinetics обсуждаются в Section 5.
 
@@
 %Содержательная постановка задачи
-
-Рассмотрим математическую модель реакции алкилирования изобутана олефинами в присутствии серной кислоты, которая представляет собой систему обыкновенных нелинейлых дифференциальных уравнений (\ref{eq:one}) -- (\ref{eq:twelve}).
-
-\begin{equation}
-  \frac{dc_1}{dt} = -k_1c_1 + k_2c_3 - k_3c_1c_3 - k_7c_1c_2c_4 - k_{11}c_1 + k_{14}c_{11}
-  \label{eq:one}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_2}{dt} = -k_4c_2c_4 - k_6c_2c_5 - k_7c_1c_2c_4 - k_{15}c_{11}c_2c_4
-  \label{eq:two}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_3}{dt} = k_1c_1 + k_4c_2c_5 - k_3(c_1 + c_{11})c_3 - k_{5}c_{12}c_3 - k_2c_3 + k_7c_1c_2c_4 + k_{15}c_{11}c_2c_4
-  \label{eq:three}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_4}{dt} = k_3(c_1 + k_{11})c_3 - k_4c_2c_4 - k_{7}c_{1}c_2c_4 - k_{15}c_{11}c_2c_4
-  \label{eq:four}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_5}{dt} = k_5c_{12}c_3 - k_{6}c_2c_5
-  \label{eq:five}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_6}{dt} = k_4c_{2}c_4
-  \label{eq:six}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_7}{dt} = k_6c_{2}c_5 - k_{10}c_7
-  \label{eq:seven}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_8}{dt} = k_7c_{1}c_2c_4 + k_{15}c_{11}c_2c_4 + k_9c_9c_{10} - k_8c_8
-  \label{eq:eight}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_9}{dt} = k_8c_{8} - k_{9}c_{9}c_{10}
-  \label{eq:nine}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_{10}}{dt} = k_8c_{8} - k_{9}c_{9}c_{10}
-  \label{eq:ten}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_{11}}{dt} = -k_3c_{11}c_3 - k_{15}c_{11}c_{2}c_4 + k_{11}c_1 + k_{12}c_{12} - k_{13}c_{11} - k_{14}c_{11}
-  \label{eq:eleven}
-\end{equation}
-
-\begin{equation}
-  \frac{dc_{12}}{dt} = -k_5c_{12}c_3 + k_{13}c_{11} - k_{12}c_{12}
-  \label{eq:twelve}
-\end{equation}
-
-The initial conditions are $t = 0, c_1 = c_1^0; c_2=c_2^0; c_3 = 0; c_4 = 0; c_5= 0; c_6 = 0; c_7 = 0; c_8 = 0; c_9 = 0; c_{10} = 0; c_{11}=c_{11}^0; c_{12} = c_{12}^0$.
-The corresponding species in eqs (\ref{eq:one}) -- (\ref{eq:twelve}) are 1, iC4H8; 2, iC4, 3, iC4+; 4, TMPs+; 5, DMHs+; 6, TMPs; 7, DMHs; 8, HEs; 9, iCx+; 10, iCy=; 11, 2-C4H8; 12, 1-C4H8.
-
-В качестве экспериментальных данных взяты данные со статьи [], которые представляют собой изменение концентраций компонентов реакции во времени при различных температурах; в качестве примера экспериментальных данных для сырья \#1 при температуре 276.2 K представлены данные в таблице \ref{table1}.
-
-\begin{table}
-\label{table1}
-\caption{Experimental Data}
-\begin{center}
-\begin{tabular}{ccccccc}
-\hline
-%\multicolumn{1}{l}{\rule{0pt}{12pt}
-%                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
-& 1 min & 2 min & 5 min & 10 min & 15 min & 20 min  \\
-\hline\rule{0pt}{12pt}
-DMH & 0.12 & 0.11 & 0.1	& 0.1 &	0.095 &	0.09  \\
-TMP & 0.54 & 0.65 & 0.69 & 0.69 & 0.7 & 0.705 \\[2pt]
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-Таким образом, решая систему (\ref{eq:one}) -- (\ref{eq:twelve}) с соответсвующими начальными данными, мы получим изменение рассчитанных концентраций компонентов реакции по времени. ...
-Необходимо минимизировать следующую целевую фукнцию:
-
-\begin{equation}
-  F = \sum ...
-  \label{eq:objective_func}
-\end{equation}
 
@@
 
-% До схемы характеристического алгоритма - переписать
-
 Функция $\varphi(y)$ предполагается многоэкстремальной и заданной в виде ``black box'' (т.е.
@@
 Такая постановка задачи полностью соответствует the inverse problem of chemical kinetics.
-
-%The Lipschitz condition (\ref{Lip}) может быть использовано для оценки глобального минимума функции на интервале, а знание константы Липшица позволяет конструировать алгоритмы глобального поиска и доказывать условия их сходимости (see, for example, \cite{Strongin2000}).
-
-%Одной из основных трудностей в решении многомерных задач глобальной оптимизации является рост вычислительных затрат при увеличении размерности задачи.
-Уменьшение числа испытаний функции при сохранении точности решения возможно за счет более полного использования априорных предположений о целевой функции, что приводит к адаптивным последовательным алгоритмам оптимизации.
 
@@
 Таким образом, search trial в некоторой точке $x'\in[0,1]$ будет включать в себя сначала the construction of the image $y'=y(x')$, и лишь затем вычисление значения функции $ z' = \varphi(y')$.
+
+\subsection{Characteristic Algorithms}
 
@@
 
-For the Piyavskii method the interval characteristic R(i) is an estimate (with the inverse sign) of the least value of the objective function $f(x)$ within the
+For the Piyavskii method the interval characteristic $R(i)$ is an estimate (with the inverse sign) of the least value of the objective function $f(x)$ within the
 interval $(x_{i-1},x_i)$. As a result, the point of a new trial is taken within the interval containing the estimate of the least value of $f(x)$ over the search domain.
@@
 
-%Связка про распараллеливание
-
 \subsection{Parallel Asynchronous Global Search Algorithm}
 
-Принадлежность GSA к классу caracteristic algorithms существенно упрощает схему его распараллеливания. 
-Как уже было отмечено, характеристика интервала $R(i)$ может рассматриваться как некоторая мера нахождения в нем интервалах точки глобального минимума. 
+Принадлежность Global Search Algorithm (GSA) к классу caracteristic algorithms подсказывает возможный путь его распараллеливания. 
+Как уже было отмечено, характеристика интервала $R(i)$ может рассматриваться как некоторая мера нахождения (локализации?) в нем точки глобального минимума. 
 Тогда вместо одного лучшего интервала можно выбрать сразу несколько интервалов с наибольшими характеристиками и провести в них очередные испытания параллельно.
-Более того, схема характеристических алгоритмов допускает и асинхронное распараллеливание, которое минимизирует простой процессоров в случае, если трудоемкость проведения испытаний зависит от конкретной точки точках области поиска.
+Более того, схема характеристических алгоритмов допускает и асинхронное распараллеливание, которое минимизирует простой процессоров в случае, если трудоемкость проведения испытаний зависит от конкретной точки области поиска.
 
@@
 
-5. Определить прообраз $x^{k+p+1} \in (x_{t-1},x_t)$ точки нового испытания в соответствии с формулой
+5. Определить preimage $x^{k+p+1} \in (x_{t-1},x_t)$ точки нового испытания в соответствии с формулой
 \[
@@
 Вещественное число $0<\epsilon<1$ и целое число $K_{max}>0$ являются параметрами алгоритма.
-
-%Описанный параллельный асинхнонный алгоритм основан на последовательном information global search algorithm.
-Теоретическое обоснование сходимости алгоритма приведено в \cite{Strongin2000}.
-Здесь же приведены схемы синхронного распараллеливания, использовавшиеся ранее при решении ряда прикладных задач \cite{Kalyulin2017,Modorskii2016}.
-%Новизна данной работы заключается в практической реализации и применении асинхронной схемы распалаллеливания, обладающей большей эффективностью при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска.
-Это подтверждается результатами экспериментов, описанными в следующем разделе.
 

### 2023-01-24 2023/problem-info/artricle.tex
Поправлен УДК
Поправлена схожесть.

@@
 
-УДК 004.222+004.272.25
+УДК 519.853.4
 
@@
 Локальный метод используется для локального уточнения и призван ускорить сходимость алгоритма.
-Искомая функция удовлетворяет условию Липшица с неизвестной константой.
-Для подтверждения теории были приведены вычислительные эксперименты, демонстрирующие ускорение при решении серии тестовых задач.
+Для целевой функции задачи оптимизации предполагается только то, что она удовлетворяет условию Липшица, при этом константа считается неизвестной.
+Для подтверждения теории был приведен ряд экспериментов на серии тестовых задач,результаты демонстрируют значительное ускорение по числу итераций алгоритма.
 
-\textbf{Ключевые слова:} глобальная оптимизация, локальная оптимизация, многоэкстремальные функции, деревья решений.
+\textbf{Ключевые слова:} деревья решений, глобальная оптимизация, многоэкстремальные функции, локальная оптимизация.
 \end{small}
@@
 
-При поиске глобального минимума функции можно придерживаться различных алгоритмов: от основанных на идее случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3}, до детерминированных алгоритмов, гарантирующих сходимость к глобальному минимуму \cite{fio_bib4, fio_bib5, fio_bib6}.
-Поскольку в реальных задачах глобальной оптимизации каждое вычисление значения функции представляет собой весьма трудоемкую задачу, необходимо уменьшить количество таких операций.
-Этого можно добиться целенаправленным выбором вариантов в процессе поиска оптимального решения.
-На этой идее основывается алгоритм глобального поиска (АГП).
-В данной работе мы попробуем объединить АГП и метод локальной оптимизации (метод Хука-Дживса) с помощью деревьев решений, с целью уменьшить количество проводимых итераций, а вместе с этим и количество вычислений целевой функции.
+При нахождении глобального минимума функции можно использовать разные подходы: от алгоритмов использующих случайный поиск \cite{fio_bib1, fio_bib2, fio_bib3}, до детерминированного, гарантирующего нахождение глобального минимума \cite{.
+fio_bib4, fio_bib5, fio_bib6}.
+В реальных задачах, связанных с глобальной оптимизацией, любое вычисление значения функции является очень трудоемкой задачей, поэтому количество таких операций должно быть уменьшено.
+Это можно сделать путем целенаправленного выбора вариантов в процессе поиска наилучшего решения.
+На этой идее основан алгоритм глобального поиска (AGS).
+В статье мы попробуем объединить АГП и метод локальной оптимизации (метод Хука-Дживса) с помощью деревьев решений, с целью уменьшить количество проводимых итераций, а вместе с этим и количество вычислений целевой функции.
 
-Алгоритмы поиска локального экстремума предназначены для определения лишь одного из локальных экстремумов на множестве допустимых решений, в котором целевая функция принимает максимальное или минимальное значение.
-При их построении могут использоваться как детерминированный спуск в область экстремума, так и случайный поиск.
-Среди детерминированных различают методы нулевого порядка и градиентные (1-го и 2-го порядка).
+Алгоритмы локальной оптимизации предназначены для нахождения одного локального экстремума в области определения задачи, в котором целевая функция принимает  минимальное (или максимальное) значение.
+При их построении может использоваться как детерминированный спуск в области экстремальных значений, так и случайный поиск.
+К детерминированным относятся методы нулевого порядка и градиентные методы (первого и второго порядка).
 
@@
 
-Рассмотрим задачу поиска глобального минимума функции $\varphi(y)$ в гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1\le\ i\le\ n \}$ При этом будем предполагать, что функция удовлетворяет условию Липшица с априори неизвестной константой $L$.
-Что соответствует ограниченности изменения значений функции при ограниченной вариации аргумента.
-Это предположение можно интерпретировать (применительно к прикладным задачам) как отражение ограниченности мощностей, порождающих изменения в моделируемой системе.
+Рассмотрим задачу глобальной оптимизации с целевой функцией $\varphi(y)$ на гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1 \le\, i\le\ n \}$ Предположим, что функция удовлетворяет условию Липшица с заранее неизвестной константой $L$.
+Условие Липшица дает нам  ограниченные вариации значений функций при ограниченных изменениях её аргументов.
+При решение прикладных задач, это предположение можно интерпретировать как указание на ограниченную способность моделируемой системы к изменениям.
 
@@
 
-Численное решение задачи (\ref{sec:problem})   сводится к построению оценки $ y_k^\ast\in\ D$ , отвечающей некоторому понятию близости к точке $y^\ast$ (например, ${||y^\ast-y}_k^\ast||\le\ \varepsilon$, где $\varepsilon\geq0$ есть заданная точность) на основе конечного числа $k$ вычислений значений оптимизируемой функции.
-Относительно класса рассматриваемых задач предполагается, что оптимизируемая функция $\varphi(y)$ может быть задана не аналитически, в виде формулы, а алгоритмически, как результат работы некоторой подпрограммы или библиотеки.
+При численном решении задачи (\ref{sec:problem}) на основе некоторого числа $k$ вычислений значений целевой функции, происходит построение оценки $y_k^\ast\in D$,  соответствующей понятию близости к точке $y^\ast$ (например, ${ ||y^\ast -y}_k^\ast||\le\ \varepsilon$, где $\varepsilon\geq0$ — заданная точность).
+Для рассматриваемого класса задач предполагается, что оптимизируемая функция $\varphi(y)$ может быть задана алгоритмически как результат работы подпрограммы или библиотеки, а не аналитически в виде формулы.
 
-Задачи многоэкстремальной оптимизации имеют существенно более высокую трудоемкость решения по сравнению с другими типами оптимизационных задач, т.к.
-глобальный оптимум является интегральной характеристикой решаемой задачи и требует исследования всей области поиска.
-Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (сетки) в области параметров, и выборе наилучшего значения функции на данной сетке.
-Вычислительные затраты на решение задачи растут экспоненциально с ростом размерности.
+По сравнению с другими типами задач оптимизации, задачи многоэкстремальной оптимизации имеют значительно более высокую сложность решения.
+Вычислительные затраты на решение задачи растут экспоненциально с увеличением размерности.
+Глобальная оптимальность является интегральным свойством решаемой задачи и требует исследования во всем пространстве поиска.
+В результате поиск глобального оптимума сводится к построению покрытия (сетки) в области параметров и выборе оптимального значения функции на этой сетке.
 
@@
 
-Основная идея данного подхода заключается в том, что минимизируемая функция рассматривается как реализация некоторого случайного процесса.
-Решающие правила алгоритма конструируются таким образом, что очередная итерация проводится в точке глобального минимума математического ожидания значений функции.
-Эта точка записывается в список известных значений и итерации повторяются.
-Так происходит до тех пор, пока не достигнут один из выбранных критериев остановки: расстояние между точками отрезка не становится меньше заданного значения или новые точки не попадают в окрестность истинного глобального минимума.
-Еще одним вариантом остановки работы алгоритма является достижение установленного максимума возможных итераций.
-\cite{fio_bib10}
+Алгоритм глобального поиска основан на идее, что минимизируемая функция представляет из себя случайный процесс.
+Решающее правило алгоритма построено так, что следующая итерация выполняется в глобальном минимуме математического ожидания значения функции.
+Найденная точка сохраняется в списке известных значений, после чего итерация повторяется .
+Это продолжается до тех пор, пока не будет достигнут один из выбранных критериев остановки:
+\begin{enumerate}
+ \item Расстояние между двумя точками не будет меньше заданного значения.
+ \item Новая точка не будут получена рядом с истинным глобальным минимумом.
+ \item Будет достигнуто заданное максимального числа возможных итераций.
+\end{enumerate}
 
@@
 сведение многомерной задачи к эквивалентной одномерной) с использованием кривых Пеано.
-Они позволяют свести многомерную задачу минимизации в области $D$  к одномерной задаче минимизации на отрезке $[0, 1]$
+Которая позволяют свести исходную многомерную задачу оптимизации в области $D$  к одномерной задаче минимизации на отрезке $[0, 1]$
 
@@
 	\end{equation}
-где $r>1$ --- заданный параметр алгоритма.
+где $r>1$ --- параметр алгоритма.
 
-\item  Для каждого интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить величину $R(i)$, называемую \textit{характеристикой} интервала, в соответствии с формулами
+\item  Для всех интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить значение $R(i)$, называемое \textit{характеристикой} интервала, в соответствии с формулами
 \begin{equation}
@@
 При решении тестовых задач, в которых точка глобального минимума $x^*$ является  известной, можно использовать также и критерий остановки по попаданию в окрестность глобального минимума.
-В этом случае метод прекращает работу, если если хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}) выполняется условие $\left|x_{t_j}-\ x^\ast\right| < \varepsilon.$
+В этом случае алгоритм останавливает свою работу, если хотя бы для одного $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}) выполняется условие $\left|x_{t_j}-\ x^\ast\right| < \varepsilon.$
 
-В качестве итоговой оценки глобально-оптимального решения рассматриваемой задачи выбираются значения 
+Результатом работы АГП при решении рассматриваемой задачи, является оценка глобального оптимума:
 \begin{equation}
@@
 
-Обоснование данного способа организации вычислений см.
-в \cite{fio_bib20}.
-Модификации, учитывающие наличие ограничений-неравенств в задаче, а также информация о производной целевой функции, представлены в \cite{fio_bib12, fio_bib9, fio_bib11}.
+В \cite{fio_bib20} приведено обоснование данного способа организации вычислений.
+А в работах  \cite{fio_bib9, fio_bib11} представлены модификации, учитывающие наличие ограничений-неравенств, а также информацию о значение производной целевой функции,
 
@@
 
-Метод Хука–Дживса --- это комбинация исследующего поиска по направлениям и поиска по образцу \cite{fio_bib14, fio_bib15}.
+ Метод Хука-Дживса представляет собой комбинацию поиска по образцу и исследующего поиска по направлениям  \cite{fio_bib14, fio_bib15}.
 
-Исследующий поиск определяется следующим образом: 
+Алгоритм исследующего поиска состоит из следующих шагов: 
 \begin{enumerate}
-\item	Определяется значение шага (оно различно для каждого направления координат и может измениться во время поиска). 
-\item	Шаг поиска считается успешным, если значение целевой функции в контрольной точке не превышает значение исходной целевой функции. 
+\item	Определяется значение шага (может меняться в процессе поиска и различно для каждой координаты).
+
+\item	Шаг поиска будет успешным, если значения целевой функции в контрольных точках не превышают исходных значений целевой функции.
+
 \item	В противном случае нужно вернуться к предыдущему пункту и сделать шаг в обратном направлении.
-\item	После перебора всех N координат исследовательский поиск заканчивается. Полученная точка называется базовой точкой.
+
+\item	После того как будут пройдены все $N$ параметров, исследующий поиск заканчивается. Найденная точка называется базовой.
 \end{enumerate}
 
-Что касается поиска по образцу, то он заключается в реализации шага из найденной в ходе исследующего поиска базовой точки вдоль прямой, соединяющей ее с предыдущей базовой точкой.
-Величина этого шага определяется задаваемым заранее параметром.
+Поиск по образцу состоит из выполнения одного шага от полученной базовой точки вдоль линии, соединяющей ее с предыдущей базовой точкой. Величина этого шага определяется задаваемым заранее параметром.
 
@@
 Процедура прогнозирования по дереву решения начинается с корня.
-От каждого не листового узла процедура идет влево или вправо на основании значения некоторой переменной, индекс которой хранится в наблюдаемом узле.
+От каждого не листового узла процедура идет вправо или влево в соответствии со значением переменной (ее индекс хранится в наблюдаемом узле).
+
 Возможны следующие переменные:
-Упорядоченные переменные, значения которых сравнивается с порогом (который также хранится в узле). Если значение меньше порога, процедура идет влево. В противном случае --- вправо.
-Категориальные переменные --- значение дискретной переменной проверяется на предмет того, принадлежит ли оно определенному подмножеству значений (также хранящемуся в узле) из набора допустимых величин.
+\begin{enumerate}
+
+\item Категориальные переменные --- значение дискретной переменной проверяется на предмет того, принадлежит ли оно определенному подмножеству значений (также хранящемуся в узле) из набора допустимых величин.
 Если принадлежит, то идем влево.
 В противном случае --- вправо.
+
+\item Упорядоченные переменные --- значения которых сравнивается с порогом (который также хранится в узле). Если значение больше порога, процедура идет вправо. В противном случае --- влево.
+\end{enumerate}
+
 Как только алгоритм достигает конечного узла, в качестве ответа процедуры прогнозирования выбирается значение, присвоенное этому узлу.
@@
 \end{equation}
-	где $r>1$ --- заданный параметр алгоритма.
+где $r>1$ --- параметр алгоритма.
 
-	\item  Для каждого интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить величину $R(i)$, называемую \textit{характеристикой} интервала, в соответствии с формулами
+\item  Для всех интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить значение $R(i)$, называемое \textit{характеристикой} интервала, в соответствии с формулами
 \begin{equation}
@@
 
-Вычислительные эксперименты проводились на персональном компьютере со следующими характеристиками: Операционная система Windows 11,  Процессор: Intel(R) Core™ i5-8250U CPU @ 1.60 GHz, 12 Gb RAM.
+Вычислительные эксперименты были проведены на компьютере со следующими характеристиками: операционная система Windows 11, процессор: Intel(R) Core™ i5-8250U CPU @ 1.60 GHz, 12 Gb RAM.
+
 В работе \cite{fio_bib13, fio_bib17} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.
 
-В таблице \ref{tab:1} отражено среднее число итераций , которые выполнил метод при решении серии задач из данных классов.
-Символ «>» отражает ситуацию, когда не все задачи класса были решены каким-либо методом.
-Количество нерешенных задач указано в скобках.
+\cite{fio_bib13, fio_bib17} описывает генераторы GKLS, которые могут генерировать многоэкстремальные задачи оптимизации с известными свойствами: глобальным минимумом и его значением, количеством локальных минимумов и так далее.
+
+В таблице \ref{tab:1} приведено среднее количество итераций методов глобальной оптимизации при решении серии задач из генератора GKLS.
+Знак «>» указывает на ситуацию, когда не все задачи были решены, в скобках указано количество нерешенных задач.
+Как видно, АГП превосходит методы DIRECT и DIRECTl по среднему числу итераций.
 
@@
 
-Как видно из таблицы 1, АГП превосходит методы DIRECT и DIRECTl на всех классах задач по среднему числу итераций.
-
-Ниже приведены результаты сравнения двух алгоритмов – стандартного АГП и его объединение с методом Хука-Дживса, при использовании деревьев решений.
+В таблице \ref{tab:2} приведены результаты сравнения двух алгоритмов –-- многомерного алгоритма глобального поиска (АГП) и АГП с использованием деревьев решений для выявления областей притяжения локальных минимумов (Деревья решений).
 Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{fio_bib19}.
-Критерием остановки служило попадание точки испытания в эпсилон окрестность истинного глобального минимума.
-При этом измерялось не только среднее число проводимых испытаний, но и среднее число точек испытаний.
-
-В таблице \ref{tab:2} представлено среднее количество итераций, и среднее число точек испытаний, проводимое двумя методами: алгоритмом глобального поиска (АГП) и объединением АГП и локального метода с использованием деревьев решений (Деревья решений) на задачах разной размерности ($N$) и разного класса (Класс задачи).
-Значение в скобках (если оно есть) показывает количество не решенных задач.
-Это означает, что было достигнут максимум по числу проводимых итераций, но в окрестность глобального минимума точка так и не попала.
+Алгоритм прекращал свою работу как только поражалась точка испытания в эпсилон окрестность истинного глобального минимума.
 
@@
 
-Итак, ускорение по числу итераций неоспоримое, причем на задачах любой размерности.
+Ускорение по числу итераций достаточно велико, причем на задачах любой размерности.
 Среднее число точек испытаний тоже, практически везде, уменьшилось, но эти значения могут быть сильно выше, чем среднее число испытаний.
@@
 
-Методы глобальной оптимизации являются областью активных научных исследований.
-В результате работы удалось успешно объединить алгоритм глобального поиска вместе с поиском локального минимума методом Хука-Дживса.
-С целью экспериментального подтверждения теоретических свойств рассматриваемого алгоритма были так же проведены вычислительные эксперименты на серии из сотни тестовых задач.
+В результате работы нами были успешно совмещены алгоритм глобального поиска с локальным поиском минимума (метод Гука-Джевса).
+Также были проведены вычислительные эксперименты на серии из сотни тестовых задач для экспериментального подтверждения теоретических свойств рассматриваемого алгоритма.
 

### 2023-01-24 2023/PAVT/ParallelDecisionTree.tex
Поправлено описание экспериментов.

@@
 
-The numerical experiments were carried out on Lobachevsky supercomputer.
-Each supercomputer node was a two-processor one (two Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM).
-All methods considered in the above sections were implemented using Globalizer software \cite{globalizerSystem} (development language – C++).
-Intel C++ 17.0.0 compiler was used to assembly the system.
+Вычислительные эксперименты были проведены на Lobachevsky supercomputer. Один узел кластера состоит из двух процессоров Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM.
 
-В работе \cite{fio_bib13, fio_bib17} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.
+\cite{fio_bib13, fio_bib17} описывает генераторы GKLS, которые могут генерировать многоэкстремальные задачи оптимизации с известными свойствами: глобальным минимумом и его значением, количеством локальных минимумов и так далее.
 
-В таблице \ref{tab:1} отражено среднее число итераций , которые выполнил метод при решении серии задач из данных классов.
-Символ «>» отражает ситуацию, когда не все задачи класса были решены каким-либо методом.
-Количество нерешенных задач указано в скобках.
+В таблице \ref{tab:1} приведено среднее количество итераций методов глобальной оптимизации при решении серии задач из генератора GKLS.
+Знак «>» указывает на ситуацию, когда не все задачи были решены, в скобках указано количество нерешенных задач.
+Как видно, АГП превосходит методы DIRECT и DIRECTl по среднему числу итераций.
 
@@
 
-Как видно из таблицы 1, АГП превосходит методы DIRECT и DIRECTl на всех классах задач по среднему числу итераций.
-
 Ниже приведены результаты сравнения двух алгоритмов – асинхронного АГП и его объединение с локальным методом (методом Хука-Дживса) при использовании decision tree.
@@
 
+В таблице \ref{tab:2} приведены результаты сравнения двух алгоритмов –-- асинхронного алгоритма глобального поиска (АГП) и АГП с использованием деревьев решений для выявления областей притяжения локальных минимумов (Деревья решений).
+Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{fio_bib19}.
+Алгоритм прекращал свою работу как только поражалась точка испытания в эпсилон окрестность истинного глобального минимума.
+В таблице сравнивается среднее количество итераций, и ускорение по времени относительно последовательного запуска.
+Распараллеливание  осуществлялось с применением асинхронного MPI алгоритма.
+При этом, запуск производился на 8 процессах, а следовательно, на каждой итерации вычислялось 7 точек.
+
 \begin{table}[h!]
@@
 		& ~ & АГПА & decision tree & АГПА & decision tree \\ \hline
-          & Simple & 2350.0 & 247.4 & 2348.0 & 390.0  \\ \hline
-        2  & Hard & 4732.3 & 353.6 & 4730.3 & 1101.1  \\ \hline
-          & Simple & 2115.5 & 579.5 & 2113.5 & 2441.4  \\ \hline
-        3  & Hard & 5347.4 & 588.1 & 5345.4 & 2532.7  \\ \hline
-          & Simple & 12168.9 & 768.2 & 12166.9 & 3582.2  \\ \hline
-        4  & Hard & 25636.5 & 960.0 & 52634.5 & 5171.6  \\ \hline
-          & Simple & 20633.5 & 978.0 & 20631.5 & 4840.4  \\ \hline
-        5  & Hard & 161094.9 & 1217.0 & 161092.9 (4) & 7115.5  \\ \hline
+		& Simple & 3823,1   & 485,6  & 7,4   & 84,1 \\ \hline
+		2  & Hard & 787,6    & 304,2  & 13,8  & 62,4  \\ \hline
+		& Simple & 5700,5   & 720,1  & 5,6   & 63,6  \\ \hline
+		3  & Hard & 2411,9   & 634,3  & 5,0   & 29,8  \\ \hline
+		& Simple & 31101,5  & 1037,4 & 5,1   & 200,7  \\ \hline
+		4  & Hard & 10418,1  & 816,2  & 6,9   & 122,7  \\ \hline
+		& Simple & 163313,5 & 1204,2 & 4,8   & 781,9  \\ \hline
+		5  & Hard & 27524,5  & 999,4  & 4,6   & 161,5  \\ \hline
 	\end{tabular}

### 2023-01-24 2023/PAVT/ParallelDecisionTree.tex
some minor corrections

@@
 
-\cite{fio_bib13, fio_bib17} описывает генераторы GKLS, которые могут генерировать многоэкстремальные задачи оптимизации с известными свойствами: глобальным минимумом и его значением, количеством локальных минимумов и так далее.
+В экспериментах мы использовали генератор тестовых задач GKLS, который может порождать многоэкстремальные задачи оптимизации с известными свойствами: точкой глобального минимума, количеством локальных минимумов, и т.п.
 
-В таблице \ref{tab:1} приведено среднее количество итераций методов глобальной оптимизации при решении серии задач из генератора GKLS.
-Знак «>» указывает на ситуацию, когда не все задачи были решены, в скобках указано количество нерешенных задач.
-Как видно, АГП превосходит методы DIRECT и DIRECTl по среднему числу итераций.
+Для демонстрации эффективности исследуемого алгоритма глобального поиска приведем результаты его сравнения с известным методами DIRECT и DIRECTl
+В таблице \ref{tab:1} приведено среднее количество итераций, выполненных соответствующими методами при решении серии задач GKLS.
+Знак «>» указывает на ситуацию, когда не все задачи были решены, здесь в скобках указано количество нерешенных задач.
+Как видно, GSA превосходит методы DIRECT и DIRECTl по среднему числу итераций, требующихся для решения задач с одинаковой точностью.
 
@@
 
-Ниже приведены результаты сравнения двух алгоритмов – асинхронного АГП и его объединение с локальным методом (методом Хука-Дживса) при использовании decision tree.
+Ниже приведены результаты сравнения двух параллельных алгоритмов --- асинхронного GSA и его модификации с использованием decision tree и локального метода Хука-Дживса.
 Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{fio_bib19}.
-Критерием остановки служило попадание trial в эпсилон окрестность истинного глобального минимума.
+Критерием остановки служило попадание точки очередного испытания trial в $\varepsilon$ окрестность истинного глобального минимума.
 
-В таблице \ref{tab:2} представлено среднее количество итераций, и ускорение по времени относительно последовательного запуска.
-Распараллеливание  осуществлялось с применением асинхронного MPI алгоритма.
-При этом, запуск производился на 4 процессах, а следовательно, на каждой итерации вычислялось 3 точки.
+В таблице \ref{tab:2} представлено среднее количество итераций и ускорение по времени относительно последовательного запуска.
+Распараллеливание  осуществлялось с применением  MPI.
+При этом запуск производился на 4 процессах, а следовательно, на каждой итерации вычислялось 3 точки.
 Значение в скобках (если оно есть) показывает количество не решенных задач.
@@
 Распараллеливание  осуществлялось с применением асинхронного MPI алгоритма.
-При этом, запуск производился на 8 процессах, а следовательно, на каждой итерации вычислялось 7 точек.
+При этом, запуск производился на 8 процессах, а следовательно, на каждой итерации вычислялось 7 точек испытаний.
 

### 2023-01-24 2023/PAVT_Ufa/PAVT2023.tex

@@
 %\title{Solving the Inverse Problems of Chemical Kinetics Using the Asynchronous Global Optimization Algorithm}
-\title{Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as Catalyst Using the Asynchronous Global Optimization Algorithm}
+\title{
+Solving ...
+}
 \titlerunning{Solving the Inverse Problems of Chemical Kinetics}  % abbreviated title (for running head)
@@
 
-The paper considers the application of parallel computing technology to the simulation of catalytic chemical reaction, which is widely used in the modern automobile industry to produce gasoline with high octane number.
+The paper considers ...
+%the application of parallel computing technology to the simulation of catalytic chemical reaction, which is widely used in the modern automobile industry to produce gasoline with high octane number.
 As a chemical reaction, the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid is assumed.
@@
 %Часть УГНТУ
-Currently, there is a tendency to improve the environmental characteristics of automobile fuel while maintaining a high octane number.
+%Currently, there is a tendency to improve the environmental characteristics of automobile fuel while maintaining a high octane number.
 Sulfuric acid alkylation of isobutane with olefins makes it possible to obtain a high-octane component of gasoline with a minimum content of aromatic hydrocarbons.
@@
 
-Так как математической моделью реакции является система дифференциальных уравнений, найти найти значения констант, входящих в него, можно только численно (см., например, \cite{RSD}).
+Так как математической моделью реакции является система дифференциальных уравнений, найти значения констант, входящих в нее, можно только численно (см., например, \cite{RSD}).
 Отметим, что в таких задачах целевая функция, как правило, является многоэкстремальной, т.е.

### 2023-01-25 2023/PAVT/ParallelDecisionTree.tex
Обновлено ускорение.

--- 2023-01-24 2023/PAVT/ParallelDecisionTree.tex
@@
 		& ~ & АГПА & decision tree & АГПА & decision tree \\ \hline
-		& Simple & 3823,1   & 485,6  & 7,4   & 84,1 \\ \hline
-		2  & Hard & 787,6    & 304,2  & 13,8  & 62,4  \\ \hline
-		& Simple & 5700,5   & 720,1  & 5,6   & 63,6  \\ \hline
-		3  & Hard & 2411,9   & 634,3  & 5,0   & 29,8  \\ \hline
-		& Simple & 31101,5  & 1037,4 & 5,1   & 200,7  \\ \hline
-		4  & Hard & 10418,1  & 816,2  & 6,9   & 122,7  \\ \hline
-		& Simple & 163313,5 & 1204,2 & 4,8   & 781,9  \\ \hline
-		5  & Hard & 27524,5  & 999,4  & 4,6   & 161,5  \\ \hline
+		& Simple & 3823,1   & 485,6  & 7,4   & 0,5 \\ \hline
+		2  & Hard & 787,6    & 304,2  & 13,8  & 1,0 \\ \hline
+		& Simple & 5700,5   & 720,1  & 5,6   & 0,1 \\ \hline
+		3  & Hard & 2411,9   & 634,3  & 5,0   & 0,1 \\ \hline
+		& Simple & 31101,5  & 1037,4 & 5,1   & 0,2 \\ \hline
+		4  & Hard & 10418,1  & 816,2  & 6,9   & 0,2 \\ \hline
+		& Simple & 163313,5 & 1204,2 & 4,8   & 1,0 \\ \hline
+		5  & Hard & 27524,5  & 999,4  & 4,6   & 0,2 \\ \hline
 	\end{tabular}

### 2023-01-25 2023/PAVT/ParallelDecisionTree.tex
Добавлена часть перевода.

@@
 %
-\title{Об использовании деревьев решений для идентификации локальных экстремумов в параллельном алгоритме глобальной оптимизации}
+\title{On the use of decision trees to identify the local extrema in parallel global optimization algorithms}
 %
@@
 \begin{abstract}
-В работе рассматривается решение задач многомерной глобальной оптимизации с применением decision tree для выявления областей притяжения локальных минимумов.
-Целевая функция задачи задана как ''черный ящик''.
-Мы лишь предполагаем, что она удовлетворяет условию Липшица с неизвестной константой. 
-Для поиска глобального минимума в задачах такого типа используется алгоритм глобального поиска.
-В настоящей работе мы предлагаем способ выделения окрестностей локальных экстремумов целевой функции на основе анализа накопленной поисковой информации.
-Проведение такого анализа средствами машинного обучения позволяет принять решение о запуске локального метода, который может ускорить сходимость алгоритма.
-Выдвинутое предположение подтверждается результатами вычислительных экспериментов, демонстрирующих ускорение при решении серии тестовых задач.
+In the present work, the solving of the multidimensional global optimization problems using decision  tree to reveal the attractor regions of the local minima is considered.
+The objective function of the  problem is defined as a 'black box''.
+We assume it only to satisfy Lipschitz condition with unknown constant.
+Global search algorithm is applied for the search of global minimum in the problems of such type.
+In the present work, we propose a method for selecting the nearness of local extrema of the objective  function based on analysis of accumulated search information.
+Conducting such an analysis using machine learning tools allows making a decision to run a local  method, which can speed up the convergence of the algorithm.
+This suggestion was confirmed by the results of numerical experiments demonstrating the speedup  when solving a series of test problems.
 % We would like to encourage you to list your keywords within
@@
 
-\section{Введение}
+\section{Introduction}
 
-В работе рассматриваются параллельные алгоритмы для решения задач многомерной глобальной оптимизации.
-Такие задачи часто возникают в случае, когда требуется выбрать значения параметров исследуемой математической модели, при которых результаты моделирования наилучшим образом соответствуют экспериментальным данным.
-Для решения задач указанного класса известен широкий спектр алгоритмов, он основанных на идее случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3} метаэвристических алгоритмов, до детерминированных алгоритмов, гарантирующих сходимость к глобальному минимуму \cite{fio_bib4, fio_bib5, fio_bib6}.
-Поскольку в реальных задачах глобальной оптимизации каждое вычисление значения функции (называемое в дальнейшем trial) представляет собой весьма трудоемкую операцию, необходимо уменьшить количество таких trials.
-Этого можно добиться целенаправленным выбором вариантов в процессе поиска оптимального решения, отсекая бесперспективные подобласти поиска и исследуя только те, в которых может содержаться решение задачи.
-На этой идее основывается global search algorithm (GSA) \cite{fio_bib7}.
-В данной работе мы попробуем объединить GSA и метод локальной оптимизации (Hooke--Jeeves method) с целью уменьшить количество проводимых испытаний.
-При этом принятие решения о запуске локального метода будет осуществляться с помощью decision tree.
+In the present work, we consider the parallel algorithms for solving the multidimensional global  optimization problems.
+Such problems arise often in the cases, when it is necessary to select the values  of parameters of the mathematical model being investigated, which the results of modeling fit best to  the experimental data at.
+For solving the problems of the class specified, many algorithms are known: from the ones based on the  idea of random search \cite{fio_bib1, fio_bib2, fio_bib3} metaheuristic algorithms to the deterministic  algorithm guarantying convergence to the global minimum \cite{fio_bib4, fio_bib5, fio_bib6}.
 
-Алгоритмы локальной оптимизации предназначены для определения лишь одного из локальных экстремумов на множестве допустимых решений, в котором целевая функция принимает экстремальное (максимальное или минимальное) значение.
-Среди локальных методов традиционно выделяют методы нулевого порядка и градиентные методы  (1-го и 2-го порядков) \cite{fio_bib8,fio_bib9}.
-Отметим, что применение методов первого порядка (не говоря уже о втором) в задачах с функциями вида ''черный ящик'' затруднено, т.к.
-здесь возникает проблема численной оценки градиента.
-Поэтому мы будем рассматривать в дальнейшем только методы нулевого порядка.
+Since in real global optimization problems each computing of the function value (hereafter called  \textit{a trial}) is a very computation-costly operation, one has to reduce the number of such trials.
+It  can be achieved by intentional choice of variants in the course of search for the optimal solution cutting  off unpromising search subdomains and investigating only these ones, which the solution of the  problem can be found in.
+The global search algorithm (GSA) is based on this idea \cite{fio_bib7}.
+In  the present work, we tried to combine GSA and a local optimization method (Hooke-Jeeves method)  to reduce the number of trials executed.
+The decision on the run of the local method will be made using  a decision tree.
 
-Для того, чтобы понять, в какой момент лучше всего запускать локальный метод в рамках глобального поиска, мы будем использовать алгоритм на основе decision tree.
-На первой стадии поиска происходит накоплении поисковой информации в виде результатов проведенных испытаний.
-Результаты испытаний мы используем для тренировки decision tree, что позволяет получить кусочно-линейную аппроксимацию целевой функции и  предсказывать на ее основе поведение целевой функции.
-Для этого бы будем сравнивать значения функции в точках, соседних к той, которую считаем подозрительной на локальный минимум.
-И на основе кусочной-линейной аппроксимации принимать решение, находится очередная точка испытания в области притяжения локального минимума или нет.
+The local optimization algorithms are intended for determining only one of local extrema within a set  of feasible solutions, which the objective function takes the extremal (maximum or minimum) value in.
+Among the local methods,  the zero-order methods and the gradient methods (of the 1\textsuperscript{st} order and of the  2\textsuperscript{nd} one) are distinguished traditionally \cite{fio_bib8,fio_bib9}.
+Note that the application of the first-order methods (not to mention the second one) in the problems  with the “black-box” type functions is difficult since the problem of numerical estimating the gradient  arises here.
+Therefore, further we will consider the zero-order methods only.
 
-Основная часть статьи имеет следующую структуру. 
-В Section \ref{SecPS} рассмотрена постановка задачи глобальной оптимизации, сделаны основные предположения о целевой функции .
-В Section \ref{SecA} приведено описание используемых методов и подходов.
-В частности, в Section \ref{SecGSA} обсуждается основная идея global search algorithm, приведены его вычислителные правила.
-Section \ref{SecHG} содержит описание используемого метода локального поиска.
-Способ построения аппроксимаций функции с использованием деревьев решений описан в Section \ref{SecDT}.
-Новый алгоритм, объединяющий глобальный и локальный поиск на основе использования деревьев решений, приведен в Section \ref{SecGSAL}.
-В следующем разделе обсуждаются особенности асинхронной схемы распараллеливания нового алгоритма.
-Section \ref{SecR} содержит результаты экспериментов, проведенных на параллельной вычислительной системе с distributed memory.
-В Заключении представлены основные выводы по проделанной работе.
+In order to determine, which moment it is best to run the local method within the framework of the  global search at we will use an algorithm based on a decision tree.
+At the first search stage, the search information is accumulated in the form of the results of the trials  executed.
+We used the trial results for training the decision tree that allows obtaining a piecewise linear  approximation of the objective function and predicting the objective function behavior on its base.
+To do so, we will compare the function values at the points neighboring to the one, which we consider  to be suspicious for a local minimum and, based on the piecewise linear approximation, make a decision  whether the next trial point falls into the attractor area of a local minimum  or not.
 
-\
+The main part of the paper has the following structure.
+In Section \ref{SecPS}, the global optimization problem statement is considered, the basic suggestions  on the objective function are made.
+In Section \ref{SecA}, the description of the methods and approaches used is given.
+In particular, in Section \ref{SecGSA},  the main idea of global search algorithm is discussed, its computational rules are given.
+Section \ref{SecHG} contains the description of the local search method used.
+The method of constructing the approximation of the function using the decision trees is described in  Section \ref{SecDT}.
+The novel algorithm combining the global search and the local one on the base of using the decision  trees is presented in Section \ref{SecGSAL}.
+In the next section,  the features of the asynchronous parallelization scheme of the new algorithm are discussed.
+Section \ref{SecR} contains the results of experiments , carried out using a parallel computer system  with distributed memory.
+In Conclusion, main conclusions on the work done are presented.
+
 \section{Problem statement}\label{SecPS}
 
-Рассмотрим задачу поиска глобального минимума функции $\varphi(y)$ в гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1\le\ i\le\ n \}$.
-При этом будем предполагать, что функция удовлетворяет the Lipschitz condition с априори неизвестной константой $L$, что соответствует ограниченности изменения значений функции при ограниченной вариации аргумента.
-Это предположение можно интерпретировать (применительно к прикладным задачам) как отражение ограниченности мощностей, порождающих изменения в моделируемой системе.
+Let us consider a problem of searching the global minimum of a function $\varphi(y)$ in a hyperinterval  $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1\le\ i\le\ n \}$.
 
@@
 \end{equation}
-где $a,b \in R$ есть заданные векторы.
+
+where $a,b \in R$ are given vectors.
+
+Also, assume the function to satisfy the Lipschitz condition with \textit{a priori} unknown constant  $L$ that corresponds to limited variation of the function values at limited variation of the argument.
+generating the changes in the system being simulated.
 
 \begin{displaymath} 
-|\varphi(y_1)-\varphi(y_2)|\leq L\parallel y_1-y_2 \parallel
-,y_1,y_2 \in D, 0<L< \infty.
+	|\varphi(y_1)-\varphi(y_2)|\leq L\parallel y_1-y_2 \parallel ,y_1,y_2 \in D, 0<L< \infty. 
 \end{displaymath}
 
-Численное решение задачи (\ref{sec:problem})   сводится к построению оценки $ y_k^\ast\in\ D$ , отвечающей некоторому понятию близости к точке $y^\ast$ (например, ${||y^\ast-y}_k^\ast||\le\ \varepsilon$, где $\varepsilon\geq0$ есть заданная точность) на основе конечного числа $k$ вычислений значений целевой функции.
-Относительно класса рассматриваемых задач предполагается, что оптимизируемая функция $\varphi(y)$ может быть задана алгоритмически, как результат работы некоторой подпрограммы или библиотеки.
+This suggestion can be interpreted (with respect to the applied problems ) as the reflection of limited  power
 
-Задачи многоэкстремальной оптимизации имеют существенно более высокую трудоемкость решения по сравнению с другими типами оптимизационных задач, т.к.
-глобальный оптимум является интегральной характеристикой решаемой задачи и требует исследования всей области поиска.
-Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (сетки) в области параметров, и выборе наилучшего значения функции на данной сетке.
-При использовании равномерных сеток вычислительные затраты на решение задачи растут экспоненциально с ростом размерности.
+The numerical solving of the problem (\ref{sec:problem})  is reduced to constructing an estimate  $ y_k^\ast\in\ D$ matching some understanding of nearness to the point $y^\ast$ (for example,  ${||y^\ast-y}_k^\ast||\le\ \varepsilon$ where $\varepsilon\geq0$ is a predefined precision) based on a  finite number $k$ of the objective function values computed.
+With respect to the class of the problems  considered, it is suggested that the optimized function $\varphi(y)$ can be defined algorithmically, as  a result of executing some subroutine or library.
 
-\section{Используемые алгоритмы}\label{SecA}
+The solving of the multiextremal optimization problems is much more computation-costly  as compared  to other types of optimization problems since the global optimum is an integral characteristic of the  problem being solved and requires investigating the whole search domain.
+As a result, the search of the  global optimum is reduced to constructing some coverage (grid) in the search domain and the choice of  the best function value on this grid.
+When using the uniform grids, the computation costs of solving the  problem grow exponentially with increasing dimensionality.
 
-Основная идея подхода к построению более экономной неравномерной сетки заключается в том, что ее точки вычисляются последовательно, а целевая функция рассматривается как реализация некоторого случайного процесса.
-Решающие правила алгоритма построения сетки конструируются таким образом, что очередная точка сетки соответствует точке глобального минимума математического ожидания значений функции.
-Эта точка записывается в список известных значений и итерации повторяются.
-Так происходит до тех пор, пока не достигнут один из выбранных критериев остановки: расстояние между точками проведенных испытаний становится меньше заданного значения или же достигается установленное максимальное число итераций.
-\cite{fio_bib10}
+\section{The algorithms used}\label{SecA}
 
-При решении многомерных задач применяется редукция размерности (т.е.
-сведение многомерной задачи к эквивалентной одномерной) с использованием кривых Пеано.
-Они позволяют свести многомерную задачу минимизации в области $D$  к одномерной задаче минимизации на отрезке $[0, 1]$
+The main idea of the approach to constructing more efficient nonuniform grid is that its points are  computed sequentially while the objective function is considered as a realization of some random  process.
+The decision rules of the grid building algorithm are constructed in such a way that the next  grid point corresponds to the global minimum point of the mathematical expectation of the function  values.
+This point is stored in the list of known values, and the iterations are repeated until one of  selected stop criteria is satisfied: either the distance between the trial points becomes less than a  predefined value or a preset maximum number of iterations is achieved \cite{fio_bib10}.
+
+When solving the multidimensional problems, the dimensionality reduction (i.e.
+the reduction of the  multidimensional problem to an equivalent one-dimensional one) using Peano curves is applied.
+These  ones allow reducing a multidimensional optimization problem in the domain $D$  to a one-dimensional  minimization problem within the interval $[0, 1]$
 
@@
 
-где функция $\varphi(y(x^\ast))$  удовлетворяют более общему H{\"o}lder condition
+where the function $\varphi(y(x^\ast))$ satisfies more general H{\"o}lder condition
 
@@
 
-Поэтому вместо исходной задачи минимизации функции $\varphi(y)$ в области $D$ можно рассматривать минимизацию одномерной функции $f(x)\ =\ \varphi(y(x))$, удовлетворяющей the H{\"o}lder condition, при $ x\in [0,1]$.
+Therefore, instead of the initial problem of minimizing the function $\varphi(y)$ in the domain $D$,  one can consider the minimization of the one-dimensional function $f(x)\ =\ \varphi(y(x))$ satisfying  the H{\"o}lder condition for $ x\in [0,1]$.
 
-\subsection{Многомерный параллельный алгоритм глобального поиска}\label{SecGSA}
+\subsection{Multidimensional parallel global search algorithm}\label{SecGSA}
 
-Основные шаги parallel global search algorithm состоят в следующем.
+The main steps of the parallel global search algorithm are as follows.
 
-На подготовительном шаге параллельно проводятся $p$ trials в произвольных внутренних точках $x^1, ...,x^p$ отрезка $[0,1]$, что соответствует первой итерации  алгоритма.
+At the preliminary step, $p$ trials are executed in parallel in arbitrary internal points $x^1, ...,x^p$ of  the interval $[0,1]$ that corresponds to the first iteration of the algorithm.
 
-Если выполнено $n \geq 1$ итераций, которым соответствуют $k=k(n)$ проведенных испытаний в точках $x^i, 1\leq i\leq k$, то точки $x^{k+1},\ldots,x^{k+p}$ поисковых испытаний следующей $(n+1)$-й итерации будут вычисляться в результате выполнения следующих действий.
+If $n \geq 1$ iterations are completed, which $k=k(n)$ executed trials in the points $x^i, 1\leq i\leq  k$ correspond to, the points $x^{k+1},\ldots,x^{k+p}$ of the search trials at the next $(n+1)^{\rm  th}$ iteration will be computed as the result of performing the following operations.
 
@@
 
-\item  Переупорядочить (нижним индексом) точки $x^i, 1\leq i\leq k$, а также граничные точки отрезка [0,1] в порядке возрастания координаты:
+	\item  Renumber (by the lower indices) the points $x^i, 1\leq i\leq k$ as well as the boundary points of  the interval [0,1] in the order of increasing coordinate:  
 	\begin{equation} 
-\label{agp1_sort}
-	0=x_0<\ x_1<\ ...\ <x_{k+1}=1.
+		\label{agp1_sort} 	0=x_0<\ x_1<\ ...\ <x_{k+1}=1. 	
 	\end{equation} 
-	и сопоставить им значения $z_i=f(x_i)$.
+	and juxtapose them with the values $z_i=f(x_i)$.
 
-\item  Вычислить текущие нижние оценки $M$ неизвестной H{\"o}lder constani $H$:
+	\item  Compute current lower estimate $M$ of the unknown H{\"o}lder constant $H$:  
 	\begin{equation} 
-\label{agp2_mu}
-	\mu=max\left\{\frac{|z_i-z_{i-1}|}{{{(x}_i-x_{i-1})}^{1/N}},\ i=1,\ldots,k\right\},\ M=\ \left\{\begin{matrix}r\mu,\ \mu>0,\\1,\ \mu=0,\\\end{matrix}\right.\
+		\label{agp2_mu} 	\mu=max\left\{\frac{|z_i-z_{i-1}|}{{{(x}_i-x_{i-1})}^{1/N}},\ i=1,\ldots,k\right\},\ M=\  \left\{\begin{matrix}r\mu,\ \mu>0,\\1,\ \mu=0,\\\end{matrix}\right.\ 	
 	\end{equation}
-где $r>1$ есть параметр алгоритма.
-Данный параметр регулирует надежность алгоритма: большие значения $r$ обеспечивают гарантированное нахождение глобального минимума, выбор меньшего значения --- ускоряет сходимость алгоритма
 
-\item  Для каждого интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить величину $R(i)$, называемую \textit{характеристикой} интервала, в соответствии с формулами
+	where $r>1$ is a parameter of algorithm.
+This parameter controls the reliability of the algorithm: the  higher values of $r$ ensure guaranteed finding of the global minimum, the choice of lower value ---  speeds up the convergence of the algorithm.
+
+	\item  For each interval $(x_{i-1},x_i), 1\leq i\leq k+1,$ compute the value $R(i)$ called \textit{a  characteristic} of the interval according to the formulae
 
 	\begin{equation} 
-\label{agp3_R1}
-R(1)=2\Delta_1-4\dfrac{z_1}{M}, \; R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
+		\label{agp3_R1} R(1)=2\Delta_1-4\dfrac{z_1}{M}, \; R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M}, 
 	\end{equation}
@@
 	\begin{equation} 
-\label{agp3_Ri}
-R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1,
+		\label{agp3_Ri} R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1, 
 	\end{equation}
-где \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
 
-\item   Упорядочить характеристики $R\left(i\right),\ 1\leq i \leq k+1,$ в порядке невозрастания 
+	where \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
+
+	\item   Arrange the characteristics $R\left(i\right),\ 1\leq i \leq k+1,$ in the order of non-increasing
+
 	\begin{equation} 
-\label{agp4_R_sort}
-	R\left(t_1\right)\geq\ R\left(t_2\right)\geq...\geq\ R\left(t_k\right)\geq\ R(t_{k+1}),\ 
+		\label{agp4_R_sort} 	R\left(t_1\right)\geq\ R\left(t_2\right)\geq...\geq\ R\left(t_k\right)\geq\ R(t_{k+1}),\  
 	\end{equation}
-и выбрать $p$ интервалов с номерами $t_j,\ 1\le\ j\le\ p$, с наибольшими значениями характеристик.
 
-\item В выбранных интервалах вычислить точки $x^{k+j},\ 1\leq j\leq p,$ в соответствии с формулами
+	and select $p$ intervals with the indices $t_j,\ 1\le\ j\le\ p$ with the highest values of characteristics.
+
+	\item Compute the points $x^{k+j},\ 1\leq j\leq p$ in the selected intervals according to the formulae
+
 \begin{equation}
@@
 
-Очередные $p$ испытаний проводятся параллельно, в вычисленных по формулам (\ref{agp5_x1}), (\ref{agp4_xi}) точках $x^{k+j},\ 1\leq j\leq p$.
-После проведения испытаний их результаты заносятся в информационную базу, и алгоритм переходит к расчету новых точек испытаний.
+The next $p$ trials are executed in parallel in the points $x^{k+j},\ 1\leq j\leq p$ computed according  to the formulae (\ref{agp5_x1}), (\ref{agp4_xi}).
+Upon completing the trials, the results of these ones  are stored in the information database, and the algorithm goes to computing new trial points.
 
-Отметим, что в прикладных оптимизационных задачах процесс проведения trial является, как правило, значительно более трудоемким по сравнению с расчетом точки испытания.
+Note that as a rule the process of executing a trial in the applied optimization problems is much more  computation-costly as compared to computing the trial point.
 
-Алгоритм останавливает свою работу в случае, если хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}) выполняется условие \(\Delta_{t_j} < \varepsilon\).
-Данный критерий остановки (наряду с обычным для итерационных методов критерием, ограничивающим число выполненных итераций) используется в прикладных задачах оптимизации, в которых точка глобального минимума $x^*$ заранее неизвестна.
+The algorithm stops in the case if the condition \(\Delta_{t_j} < \varepsilon\) is satisfied for even a  single value $t_j,\ 1\le\ j\le\ p$ from (\ref{agp4_R_sort}).
+This stop criterion (along with the criterion limiting the number of executed iterations usual for the  iteration methods) is used in the applied optimization problems, which the global minimum point  $x^*$ is unknown \textit{a priori} in.
 
-При решении тестовых задач, в которых точка глобального минимума $x^*$ является  известной, можно использовать также и критерий остановки по попаданию в окрестность глобального минимума.
-В этом случае метод прекращает работу, если хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}) выполняется условие $\left|x_{t_j}-\ x^\ast\right| < \varepsilon.$
+When solving the test problems, which the global minimum point $x^*$ is known in one can use the  stop criterion upon falling into the nearness of the global minimum as well.
+In this case, the method  stops if the condition $\left|x_{t_j}-\ x^\ast\right| < \varepsilon$ is satisfied even for one value of $t_j,\  1\le\ j\le\ p$ from (\ref{agp4_R_sort}).
 
-В качестве итоговой оценки глобально-оптимального решения рассматриваемой задачи выбираются значения 
+As the final estimate of the globally optimized solution of the considered problem, the values
+
 \begin{equation} 
@@
 
-Обоснование данного способа организации вычислений см.
-в \cite{fio_bib20}.
-Модификации, учитывающие наличие ограничений-неравенств в задаче, а также информация о производной целевой функции, представлены в \cite{fio_bib12, fio_bib9, fio_bib11}.
+are taken.
+
+The substantiation of this method of organization of computations see in \cite{fio_bib20}.
+The  modifications taking into account the presence of  constraints (inequalities) in the problem as well as  the information on the derivative of the objective function  are presented in \cite{fio_bib12, fio_bib9,  fio_bib11}.
 
@@
 
-Метод Хука–Дживса относится к классу методов нулевого порядка.
-Его вычислительные правила представляет собой комбинацию исследующего поиска (для выбора направления) и поиска по выбранному направлению \cite{fio_bib14, fio_bib15}.
+The Hooke--Jeeves method belongs to the class of zero-order methods.
+Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{fio_bib14, fio_bib15}.
 
-Исследующий поиск выполняется следующим образом: 
-\begin{enumerate}
-\item	Определяется значение шага (оно различно для каждой координаты и может измениться во время поиска). 
-\item	Шаг поиска считается успешным, если значение целевой функции в контрольной точке не превышает значение целевой функции в исходной точке. 
-\item	В противном случае нужно вернуться к предыдущему пункту и сделать шаг в обратном направлении. 
-\item	После перебора всех $N$ координат исследующий поиск заканчивается. Полученная точка называется базовой точкой.
+The investigating search is performed as follows:  \begin{enumerate} \item	The step value is determined (it is different for each coordinate and can be varied in the course  of search).
+\item	The search step is considered to be successful if the value of the objective function in the check  point doesn't exceed the value of the objective function in the initial point.
+\item	Otherwise, it is necessary to return to the previous item and make a step in the reverse direction.
+\item	After making the steps in all $N$ coordinates, the investigating search is completed.
+The point  obtained is called the base point.
 \end{enumerate}
 
-Что касается поиска по направлению, то он заключается в реализации шага из найденной в ходе исследующего поиска базовой точки вдоль прямой, соединяющей ее с предыдущей базовой точкой.
-Величина этого шага определяется задаваемым заранее параметром.
+With regard to the search in a direction, it consists in performing a step from the base point found  during the investigating search along the straight line direction connecting this one with the previous  base point.
+The magnitude of this step is defined by a parameter set in advance.
 
@@
 			\includegraphics[width=1\linewidth]{figure/fig1.png} 
-			\caption{Пример итераций алгоритма Хука-Дживса} %% подпись к рисунку
+			\caption{An example of iterations of the Hooke-Jeeves algorithm} %%  подпись к рисунку 
 			\label{fig:fig1} 
@@
 
-Дерево принятия решений (decision tree) --- это средство, используемое для автоматического анализа больших массивов данных, который применяется в машинном обучении.
-Decision tree представляет собой бинарное дерево (дерево, в котором каждый не листовой узел имеет два дочерних узла).
-Его можно использовать как для решения задач классификации, так и для задач регрессии.
-При решении задач классификации каждый лист дерева помечается меткой класса, при этом несколько листьев могут иметь одну и ту же метку.
-В случае построения регрессии каждому листу дерева присваивается константа, поэтому получаемая функция аппроксимации является кусочно-постоянной.
-В нашем случае decision tree строит функцию $\varphi(y)$ кусочно постоянную аппроксимацию$\varphi(y)$  в многомерном пространстве, обозначим как $z' = \psi(y)$ значение вычисленное по decision tree в точке $y$.
+The decision tree is a tool used for an automated analysis of big data arrays, which is applied in the  machine learning.
+Decision tree is a binary tree (a tree, which each non-leaf node has two child nodes  in).
+It can be used for solving the classification problems as well as the regression ones.
+When solving  the classification problems, each leaf of the tree is marked by a class label, and several leafs can have  the same labels.
+In the case of constructing a regression, a constant is assigned to each leaf of the tree.
+Therefore, the approximating function obtained is a piecewise-constant one.
+In our case, the decision tree constructs the function $\varphi(y)$ as a piecewise-constant  approximation $\varphi(y)$  in multidimensional space.
+Let us denote the value  computed by the  decision tree in the point $y$ as $z' = \psi(y)$.
 
-При реализации алгоритма  решения задач многомерной глобальной оптимизации с применением decision tree для выявления областей притяжения локальных минимумов нами были использованы алгоритмы из библиотеки OpenCV.
-OpenCV --- библиотека алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения с открытым кодом.
-Подробнее с decision tree можно ознакомиться в \cite{fio_bib16}
+When implementing the algorithm for solving the multidimensional global optimization problems using  decision tree, we used the algorithms from OpenCV library to find the attractor areas of the local  minima.
+OpenCV is a library of the algorithms of computer vision, image processing, and  general-purpose numerical algorithms with open code.
+More details on the decision tree can be found  in \cite{fio_bib16}.
 
-\section{Объединение локального метода оптимизации и GSA для решения многомерных задач}\label{SecGSAL}
+\section{Combination of local optimization method and GSA for solving multidimensional  problems}\label{SecGSAL}
 
-В текущем разделе приведем подробное описание того, как мы использовали decision tree для выявления областей притяжения локальных экстремумов. 
-В процессе работы GSA необходимо определить, стоит ли использовать текущую точку в качестве стартовой для локального метода или нет.
-Для этого можно проверить точки соседних испытаний и если среди них нет ни одной такой, значение функции в которой меньше, чем в текущей, то можно предположить, что мы находимся в области притяжения локального минимума.
-В этом случае из текущей точки можно запустить локальный метод, который быстро сойдется к локальному минимуму функции.
-Сделать это можно только в многомерном пространстве, а не на одномерном отрезке после редукции размерности.  
-Во-первых, редуцированная функция $\varphi(y(x))$ меняет свои свойства: один локальный минимум в многомерном пространстве может разделиться на множество минимумов после редукции размерности.
-Во-вторых, после использования отображения $y(x)$ мы можем потерять информацию о взаимном расположении точек в исходном пространстве.
-Точки, расположенные близко в многомерном пространстве, могут оказаться существенно разнесенными на одномерном отрезке.
+In the current section, let us present a detailed description how we used the decision tree for finding  the attractor areas of the local extrema.
+In the course of running GSA, it is necessary to determine whether it is worth to use current point as a  start one for the local method or not.
+To do so, one can check the points of adjacent trials.
+If among  these ones there are no points such that the function values in which are lower than in current one, one  can suggest that we are in the attractor of a local minimum.
+In this case, one can run the local method  from current point, which will converge to a local minimum of the function rapidly.
+One can do it in the multidimensional space only, not in one-dimensional interval after the  dimensionality reduction.
+First, the reduced function $\varphi(y(x))$ changes its properties: one local minimum in the  multidimensional space may divide into a set of minima after the dimensionality reduction.
+Second, after the use of the mapping $y(x)$, we can lose the information on the mutual arrangement of  the points in the original space.
+The points located close to each other in the multidimensional space  may appear to be separated essentially in the one-dimensional interval.
 
@@
 			\includegraphics[width=1\linewidth]{figure/fig2.png} 
-			\caption{Функция из генератора gkls до использования кривых Пеано} %% подпись к рисунку
+			\caption{A function from the GKLS generator before using Peano curves} %%  подпись к рисунку 
 			\label{fig:fig2} 
@@
 			\includegraphics[width=1\linewidth]{figure/fig2.png} 
-			\caption{Рисунок с двумерным decision tree} %% подпись к рисунку
+			\caption{A figure with a two-dimensional decision tree} %% подпись к  рисунку 
 			\label{fig:fig2_2} 
@@
 
-Поэтому определять области притяжения локальных экстремумов мы будем в исходном многомерном пространстве.
-Однако определить соседние точки в многомерном пространстве достаточно трудоемко, и для этого мы будем использовать decision tree.
+Therefore, we will determine the attractor areas of the local extrema in the original multidimensional  space.
+However, the determining of the adjacent points in the multidimensional space is  computational-costly enough.
+To do so, we will use the decision tree.
 
-После проведения определенного числа trial (например, $100\ \ast\ N$), используем все накопившиеся точки для инициализации подходящей структуры данных и тренировки на ее основе decision tree.
-Чтобы проще было определить соседние точки, мы построим равномерную сетку с определенным шагом по каждой из координат, после чего вычислим значения аппроксимации целевой функции в этих точках с использованием дерева решений.
-Теперь, имея равномерную сетку точек и зная значение функции в каждой из них, найдем ближайшую, с точки зрения евклидова расстояния, к исходной.
-Эта ближайшая точка является проекцией исходной точки на равномерную сетку, на которой можно легко определить ее соседей.
-А так как decision tree строит аппроксимацию исходной функции (пусть и кусочно постоянную), то мы можем оценить значение функции в тех областях, в которых испытания ранее не проводились.
+After completing a definite number of trials (for example, $100\ \ast\ N$), let us use all accumulated  points to initialize a suitable data structure and training the decision tree on its base.
+To make the  determining the adjacent points easier, we construct a uniform grid with certain step in each coordinate.
+Afterwards, we compute the values of the approximation of the objective function in these points  using the decision tree.
+Now, having a uniform grid of points and knowing the function values in each  point, let us find the closest point to the initial one (from the viewpoint of the Euclidian distance).
+This  closest point is a projection of the initial point onto the uniform grid, which its neighbors can be  determined easily on.
+Since the decision tree constructs the approximation of the initial function (even  piecewise constant), we can estimate the function values in the areas, which the trials haven’t been  executed earlier in.
 
-При рассмотрении соседей необходимо учитывать следующее: если хоть один из соседей имеет значение функции меньшее, чем в текущей точке, то запуск локального метода не нужен.
-Если хотя бы в одной из соседних точек значение функции такое же, как в текущей, то ее также необходимо проверить (т.е.
-найти всех ее соседей и проверять их аналогично).
-И только в том случае, если все значения функции в соседних точках больше, мы можем запускать локальный метод.
+When considering the neighbors, it is necessary to take into account the following.
+If even a single  neighbor has a lower function value than current point, there is no need to run the local method.
+If in  even a single neighboring point the function value is the same as in the current point, it is also necessary  to check (i.
+e.
+to find all its neighbors and to heck these ones in the same way).
+Only in the case when  all function values in the neighboring points are greater than in the current point, we can run the local  method.
 
-На Fig. \ref{fig:fig2} отображены линии уровня целевой функции с отображением точек испытаний, а на Fig. \ref{fig:fig2_2} отображено соответствующее decision tree.
+Fig. \ref{fig:fig2} presents a contour plot of an objective function and the trial points; Fig.  \ref{fig:fig2_2} presents corresponding decision tree.
 
-\section{Асинхронная схема распараллеливания алгоритма}\label{SecASP}
+\section{Asynchronous algorithm parallelization scheme}\label{SecASP}
 
-Способ организации параллельных вычислений, описанный в предыдущем разделе, подразумевает синхронное проведение $p$ поисковых испытаний в точках, вычисленных на текущей итерации.
-При этом переход к следующей итерации будет выполнен лишь после окончания проведения всех испытаний.
-В случае различной трудоемкости проведения испытаний в разных точках области поиска данный подход может привести к дисбалансу вычислительной нагрузки. 
-Исправить этот недостаток можно введением асинхронности.
-Идея асинхронной схемы состоит в том, чтобы не дожидаться окончания проведения всех $p$ испытаний, а инициировать проведение нового испытания сразу же, как только было закончено хотя бы  одно их них.
+The method of parallel computing organization described in  previous section implies synchronous  performing of $p$ search trials in the points computed at current iteration.
+The next iteration will be  started upon completing all trials only.
+In the case of different computation costs of executing the trials in different points of the search  domain, this approach may lead to a misbalance of the computing jobs.
+This drawback can be corrected by introduction of asynchrony.
+The idea of the asynchronous scheme is not to wait for completing all $p$ trials but to initiate the  execution of a new trial immediately upon completing at one trial.
 
-Предположим, что главный процесс вычисляет одну точку следующего trial на каждой итерации и отправляет ее рабочему процессу для выполнения trial.
-Отметим, что в прикладных оптимизационных задачах время выполнение trial рабочим процессом -- гораздо более затратная в вычислительном отношении операция, чем выбор мастером точки нового trial.
-Тем самым простой рабочих процессов будет пренебрежимо мал.
-В этом случае (в отличие от синхронных параллельных алгоритмов) общее количество trial, выполняемых каждым рабочим процессом, будет зависеть от вычислительных затрат на выполнение конкретного trial и не может быть оценено заранее.
-При описании параллельного алгоритма предположим, что в нашем распоряжении имеется $p+1$ вычислительных процессов: один главный и $p$ рабочих.
+Let us assume that the master process computes one point of the next trial at every iteration and sends  it to the performing process to execute the trial.
+Note that executing a trial by the performing process in  the applied optimization problems is much more computation costly than the choice of a new trial point  by the master process.
+Therefore, the idle time of the performing processes will be negligible.
+In this  case (unlike the synchronous parallel algorithms) total number of trials executed by each performing  process will depend on the computation costs for executing particular trials and cannot be estimated in  advance.
+When describing the parallel algorithm, let us assume that we have $p+1$ computational  processes at our disposal: one master process and $p$ performing ones.
 
- В начале поиска главный процесс (допустим, это процесс с идентификатором $0$) инициирует параллельное выполнение $p$ trial в $p$ разных точках области поиска.
+In the beginning of search, the master process (assume it to have the identifier $0$) initiates parallel  execution of $p$ trials in $p$ different points of the search domain.
 
-Две из этих точек являются граничными, а остальные -- внутренними, т.е. в точках $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$, где
-$x^1=,x^p=,x^i\in\left(0,1\right),i=2,\ldots,p$.
+Two of these points are the boundary ones, the rest are the internal points i.
+e.
+in the points  $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$ where $x^1=,x^p=,x^i\in\left(0,1\right),i=2,\ldots,p$.
 
-Теперь предположим, что выполнено $k\geq0$ trial и рабочие процессы выполняют trial в точках $y\left(x^{k+1}\right),\;y\left(x^{k+2}\right),\ldots,\;y\left(x^{k+p}\right)$.
+Now assume $k\geq0$ trials to be completed, and the performing processes are performing the trials in  the points $y\left(x^{k+1}\right),\;y\left(x^{k+2}\right),\ldots,\;y\left(x^{k+p}\right)$.
 
-Каждый рабочий процесс, завершив в какой-то момент вычисление функции, отправляет главному процессу вычисленное значение.
-В свою очередь, главный процесс выбирает новую точку $x^{k+p+1}$ для рабочего процесса в соответствии с описанными ниже правилами.
-Заметим, что в этом случае мы будем иметь набор preimages of the trial points
-$I_k=\left\{x^{k+1},x^{k+2},\ldots,x^{k+p}\right\}$
-в которых испытания уже начаты, но еще не завершены.
+Each performing process having completed the computing of the function at some time moment sends  the computed value to the master process.
+In turn, the master process selects a new point  $x^{k+p+1}$ for the performing process according to the rules described below.
+Note that in this case  we will have a set of preimages of the trial points  $I_k=\left\{x^{k+1},x^{k+2},\ldots,x^{k+p}\right\}$ which the trials have been started already but haven’t been completed yet in.
 
-Таким образом параллельный асинхронный алгоритм глобальной оптимизации с использованием decision tree для выявления локальных областей состоит из следующих шагов:
+So far, the parallel asynchronous global optimization algorithm using decision tree to find the local  extrema attractors consists of the following steps:
 
-\begin{enumerate}
-\item  Упорядочить нижним индексом (в порядке возрастания) множество preimages of the trial points
+\begin{enumerate} \item Arrange n the increasing order (by the lower indices) the set of preimages of the trial points
 
@@
 
-содержащие все точки, в которых либо завершились, либо идут trial, т.е. получить упорядоченный набор
+	containing all points, which the trials either have been completed or are being executed in i. e. get an  ordered set
 
@@
 
-\item  Вычислить значения
+	\item Compute the values
 
@@
 
-где $ z_i=\varphi\left(y\left(x_i\right)\right)$, если $x_i\notin I_k,\;1\le i\le k+p$.
-Значения $z_i$ в точках $x_i\in I_k$ не определены, так как trial в точках $x_i\in I_k$ еще не завершены.
-Если значение $M$ равно $0$, то установить $M=1$.
+	where $ z_i=\varphi\left(y\left(x_i\right)\right)$ if $x_i\notin I_k,\;1\le i\le k+p$.
+The values $z_i$ in  the points $x_i\in I_k$ are undefined since the trials in the points $x_i\in I_k$ haven’t been completed  yet.
+If the value of $M$ equals to $0$, set $M=1$.
 
-\item  Сопоставить каждыму интервалу $\left(x_{i-1},x_i\right),\;x_{i-1}\notin I_k,x_i\notin I_k,\;2\le i\le k+p$ его характеристику $R\left(i\right)$, вычисляемую по формуле
+	\item Juxtapose each interval $\left(x_{i-1},x_i\right),\;x_{i-1}\notin I_k,x_i\notin I_k,\;2\le i\le  k+p$ with its characteristic $R\left(i\right)$ computed according to the formula
 
@@
 
-где $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$ и $ r>1$ -- параметр надежности метода.
+	where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$ and $ r>1$ is the reliability parameter of the method.
 
-\item  Выбрать интервал $\left[x_{t-1},x_t\right]$, которому соответствует максимальная характеристика, т.е.
+	\item Select the interval $\left[x_{t-1},x_t\right]$, which the highest characteristics correspond to, i. e.
 
@@
 
-\item  Определить новую точку испытания $y^{k+p+1}=y\left(x^{k+p+1}\right)$, прообразом которой является $x^{k+p+1}\in\left(x_{t-1},x_t\right)$, по формуле
+	\item Define a new trial point $y^{k+p+1}=y\left(x^{k+p+1}\right)$, the preimage of which is  $x^{k+p+1}\in\left(x_{t-1},x_t\right)$ according to the formula
 
@@
 
-\item  Получить вычисленное  значение функции с процесса $j$, добавить новый trial $z_j = f(y(x_j))$, к множеству $V$.
+	\item Get the computed function value from the process $j$, add new trial $z_j = f(y(x_j))$ to the set  $V$.
 
@@
 
-\item 	Если $ k\ <\ 100\ast\ N$, то вернуться к шагу 1.
+	\item If $k\ <\ 100\ast\ N$, return to Step 1.
 
-%\item 	Если используем decision tree не впервые, то перейти к пункту 15.
+	%\item  If the decision tree is used not in the first time, go to Step 15.
 
-\item Создаем decision tree по множеству $I_k$, получаем функцию аппроксимации  $\psi(y)$.
+	\item Create the decision tree using the set $I_k$, get the approximating function $\psi(y)$.
 
-\item 	Если используем decision tree впервые, то строим равномерную сетку
+	\item If the decision tree is used for the first time, construct a uniform grid
 
@@
 
-\item 	Вычисляем значения аппроксимации: $Z' = \{ z'=  \psi(y'), y' \in Y'\}$
+	\item Compute the approximation values: $Z' = \{ z'=  \psi(y'), y' \in Y'\}$
 
-\item Для всех точек $y'\in V$:
+	\item For all points $y'\in V$:
 
-Находим точку $y'_q$ ближайшую  к $y'$,
+	Find the points $y'_q$ closest to $y'$,
 
-Производим обход соседей $y'_q$ по описанному ранее принципу.
+	Make a detour of the neighbors $y'_q$ according to the principle described above.
 
-Если ни у одной соседней точки не нашлось значений функции меньших, чем в $y'_q$, то запускаем локальный метод.
+	If no one neighbor point has a lower value than in $y'_q$, run the local method.
 
-Очищаем множество $V$.
+	Clear the set $V$.
 
@@
 
-При вычислении точки очередного испытания главный процесс добавляет ее к множеству $I_k$ и отправляет рабочему процессу, который инициирует проведение испытания в ней.
-Главный процесс завершает алгоритм, если выполняется одно из двух условий: $\Delta_t<$ или $k+p>K_{max}$.
-Вещественное число $\epsilon>0$ и целое число $K_{max}>0$ являются параметрами алгоритма и соответствуют точности поиска решения и максимальному количеству испытаний соответственно.
+Upon computing the next trial point, the master process adds it to the set $I_k$ and sends to the  performing process, which initiates the new trial in this point.
+The master process stops the algorithm, if one of the two conditions is satisfied: $\Delta_t<$ or  $k+p>K_{max}$.
+The real value $\epsilon>0$ and the integer value $K_{max}>0$ are the parameters of the algorithm  and correspond to the precision of the search for the solution and the maximum number of trials,  respectively.
 
@@
 	\begin{center} 
-		\begin{minipage}[h]{0.9\linewidth}
-			\includegraphics[width=1\linewidth]{figure/fig3.png}
-			\caption{Блок-схема объединения АГП и метода Хука-Дживса с использованием decision tree} %% подпись к рисунку
-			\label{fig:fig3}
+				\begin{minipage}[h]{0.9\linewidth} 			\includegraphics[width=1\linewidth]{figure/fig3.png}
+					 			\caption{Block scheme of combined GSA and Hooke-Jeeves method using the  decision tree} %% подпись к рисунку 			\label{fig:fig3}
 				 \end{minipage}

### 2023-01-25 2023/PAVT/ParallelDecisionTree.tex
fix Result section

--- 2023-01-25 2023/PAVT/ParallelDecisionTree.tex
@@
 
-Ниже приведены результаты сравнения двух параллельных алгоритмов --- асинхронного GSA и его модификации с использованием decision tree и локального метода Хука-Дживса.
+%Ниже приведены результаты сравнения двух параллельных алгоритмов --- асинхронного GSA и его модификации с использованием decision tree и локального метода Хука-Дживса.
 Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{fio_bib19}.
 Критерием остановки служило попадание точки очередного испытания trial в $\varepsilon$ окрестность истинного глобального минимума.
-
-В таблице \ref{tab:2} представлено среднее количество итераций и ускорение по времени относительно последовательного запуска.
+%В таблице \ref{tab:2} представлено среднее количество итераций и ускорение по времени относительно последовательного запуска.
 Распараллеливание  осуществлялось с применением  MPI.
@@
 
-В таблице \ref{tab:2} приведены результаты сравнения двух алгоритмов –-- асинхронного алгоритма глобального поиска (АГП) и АГП с использованием деревьев решений для выявления областей притяжения локальных минимумов (Деревья решений).
-Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{fio_bib19}.
-Алгоритм прекращал свою работу как только поражалась точка испытания в эпсилон окрестность истинного глобального минимума.
-В таблице сравнивается среднее количество итераций, и ускорение по времени относительно последовательного запуска.
-Распараллеливание  осуществлялось с применением асинхронного MPI алгоритма.
-При этом, запуск производился на 8 процессах, а следовательно, на каждой итерации вычислялось 7 точек испытаний.
+Ниже приведены результаты сравнения двух параллельных алгоритмов --- асинхронного global search algorithm (GSA) и его модификации с использованием decision tree для выявления областей притяжения локальных минимумов (Decision Tree).
+Численное сравнение проводилось на классах функций GKLS двух видов (Simple and Hard) размерности 2, 3, 4 и 5 из \cite{fio_bib19}.
+Критерием остановки алгоритмов служило попадание точки очередного испытания в $\varepsilon$-окрестность истинного глобального минимума.
+В таблице сравнивается среднее количество итераций, выполненных алгоритмами при решении задач, и ускорение по времени относительно последовательного запуска.
+Распараллеливание  осуществлялось с применением технологии MPI, запуск производился на 8 процессах.
 
@@
 
-\section{Заключение}\label{SecC}
+%\section{Заключение}\label{SecC}
 
-Методы глобальной оптимизации являются областью активных научных исследований.
-В результате работы удалось успешно объединить алгоритм глобального поиска вместе с поиском локального минимума методом Хука-Дживса.
-С целью экспериментального подтверждения теоретических свойств рассматриваемого алгоритма были так же проведены вычислительные эксперименты на серии из сотни тестовых задач.
+%Методы глобальной оптимизации являются областью активных научных исследований
 
-При использовании в качестве определяющего правила о необходимости запуска локального метода decision tree, нам удалось получить достаточно хорошее ускорение.
-Применяя параллельную версию Алгоритма Глобального поиска ускорение удалось сохранить, что является несомненным преимуществом.
-Таким образом, наша схема позволяет получить преимущество от обоих из подходов (и от распараллеливания, и от запуска локального метода).
+Таким образом, в результате проведенной работы удалось успешно объединить алгоритм глобального поиска с методом локальной оптимизации. 
+В отличие от известных мультистартовых схем решение о запуске локального метода производится с использованием деревьев решений.
+Использование такой комбинации методов позволяет значительно ускорить работу алгоритма.
+
+Параллельный вариант алгоритма сохраняет свойства своего последовательного прототипа, что подтверждается вычислительными экспериментами, во время которых была решена серия из нескольких сотен задач разной размерности.
+Предложенная нами схема позволяет использовать преимущества обоих подходов -- и распараллеливания, и быстрого поиска локальных экстремумов.
 

### 2023-01-26 2023/PAVT_Ufa/PAVT2023.tex
add experiment discription

@@
 
+Предварительные эксперименты позволили оценить примерное время одного вычисления целевой функции при использовании последовательного алгоритма.
+Эксперименты проводились на архитектуре Intel(R) Core(TM) i7-10750H CPU @ 2.60GHz 2.59 GHz.
+Оценка одного вычисления целевой функции составила $195$ мс, что характеризует задачу как вычислительно сложную.
+
+Для проведениях полноценного эксперимента на $1 000 000$ итераций было осуществлено распараллеливание с помощью MPI.
+В силу сложности задачи использовался критерий остановки по числу итераций.
+Если бы такое число испытаний проводилось последовательным алгоритмом, то примерное время решения задачи составило бы $54$ часа.
+
+Численные эксперименты проводились при помощи параллельного асинхронного алгоритма, изложенного в Section 4, с использованием $160$ процессов ($159$ процессов вычисляли значения целевой функции, $1$ процесс управлял алгоритмом).
+Для проведения экспериментов использовался UNN supercomputer Lobachevsky (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node).
+Асинхронный алгоритм глобальной оптмизации был реализован на C++ (GCC 9.5.0 and Intel MPI were used); вычисление значений целевой функции было реализовано в Python 3.9.
+
+Было получено лучшее решение $0.540750$ при параметре регуляризации $alpha = 0.001$. Время решения составило около $14$ минут. Ускорение составило примерно $230$ раз.
+
+Существенное ускорение при использовании параллельного алгоритма позволило также провести исследование зависимости решения от параметра регуляризации $alpha$.
+В таблице \ref{table_1} приведены минимальные значения objective function, найденные при соответсвующем значении параметра регуляризации alpha, время решения задачи (в секундах) и ускорение по времени по сравнению с оценкой времени решения задачи последовательным методом.
+При запусках были использованы параметр метода $r=4.0$ из (\ref{R}) и точность $\epsilon = 10^{-4}\left\|b-a\right\|$.
+После выполнения заданного числа итераций или достижения методом глобального поиска заданной точности решение уточнялось Hook-Jeeves local method \cite{HookJeeves} с точностью $\epsilon = 10^{-4}\left\|b-a\right\|$.
+
+\begin{table}
+\caption{Исследование влияния параметра регуляризации alpha на минимальное значение функции}
+\label{table_1}
+\begin{center}
+\begin{tabular}{cccc}
+\hline\noalign{\smallskip}
+ Alpha      & Minimum  & Time (sec.) & Speedup \\
+\hline\noalign{\smallskip}
+0.1		&	0.810454	&	806.192	&	241       \\
+0.09	&	0.837868	&	828.649	&	234       \\
+0.08	&	0.808238	&	811.852	&	239       \\
+0.07	&	0.759134	&	803.626	&	242       \\
+0.06	&	0.723404	&	810.9	&	240       \\
+0.05	&	0.729415	&	804.674	&	242       \\
+0.04	&	0.736035	&	816.387	&	238       \\
+0.03	&	0.68152		&	809.185	&	240       \\
+0.02	&	0.669092	&	155.837	&	3         \\
+0.01	&	0.632944	&	825.238	&	235       \\
+0.009	&	0.63001		&	814.48	&	239       \\
+0.008	&	0.604817	&	830.448	&	234       \\
+0.007	&	0.619546	&	807.335	&	241       \\
+0.006	&	0.612614	&	834.703	&	233       \\
+0.005	&	0.615638	&	815.45	&	238       \\
+0.004	&	0.579725	&	821.421	&	237       \\
+0.003	&	0.583529	&	831.454	&	234       \\
+0.002	&	0.547907	&	850.017	&	229       \\
+0.001	&	0.54075		&	853.898	&	227       \\
+\noalign{\smallskip}\hline
+\end{tabular}\end{center}\end{table}
+
+Детальное исследование области значений параметра регуляризации в окрестности alpha равного нулю (от 0.1 до 0.01 с шагом 0.01 и от 0.01 до 0.001 с шагом 0.001) выявило убывание значения минимума с уменьшением параметра $alpha$, см.
+рис.
+...
+.
+
 \section{Conclusions and Future Work}

### 2023-01-26 2023/PAVT_Ufa/PAVT2023.tex
results in Russian

@@
 Эксперименты проводились на архитектуре Intel(R) Core(TM) i7-10750H CPU @ 2.60GHz 2.59 GHz.
-Оценка одного вычисления целевой функции составила $195$ мс, что характеризует задачу как вычислительно сложную.
-
-Для проведениях полноценного эксперимента на $1 000 000$ итераций было осуществлено распараллеливание с помощью MPI.
-В силу сложности задачи использовался критерий остановки по числу итераций.
-Если бы такое число испытаний проводилось последовательным алгоритмом, то примерное время решения задачи составило бы $54$ часа.
+Среднее время вычисления целевой функции составило порядка $0.2$ сек., что характеризует задачу как вычислительно сложную.
+Решение задачи последовательным алгоритмом потребовало бы не менее 1 млн.
+испытаний и, соответственно, не менее 54 часов (оценочное значение).
+Поэтому для проведения расчетов использовался только параллельный алгоритм с критерием остановки по числу испытаний $K_{max}=10^6$.
 
@@
 Для проведения экспериментов использовался UNN supercomputer Lobachevsky (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node).
-Асинхронный алгоритм глобальной оптмизации был реализован на C++ (GCC 9.5.0 and Intel MPI were used); вычисление значений целевой функции было реализовано в Python 3.9.
+Асинхронный алгоритм глобальной оптимизации был реализован на C++ (GCC 9.5.0 and Intel MPI were used); вычисление значений целевой функции было реализовано в Python 3.9.
 
-Было получено лучшее решение $0.540750$ при параметре регуляризации $alpha = 0.001$. Время решения составило около $14$ минут. Ускорение составило примерно $230$ раз.
-
-Существенное ускорение при использовании параллельного алгоритма позволило также провести исследование зависимости решения от параметра регуляризации $alpha$.
-В таблице \ref{table_1} приведены минимальные значения objective function, найденные при соответсвующем значении параметра регуляризации alpha, время решения задачи (в секундах) и ускорение по времени по сравнению с оценкой времени решения задачи последовательным методом.
+%Ускорение составило примерно $230$ раз.
+Малое время решения задачи при использовании параллельного алгоритма (примерно 14 минут на 1 задачу) позволило провести полноценное исследование зависимости решения от параметра регуляризации $\alpha$.
+В таблице \ref{table_1} приведены минимальные значения objective function, найденные при соответствующем значении параметра регуляризации, время решения задачи (в секундах) и time speedup по сравнению с оценкой времени решения задачи последовательным методом.
 При запусках были использованы параметр метода $r=4.0$ из (\ref{R}) и точность $\epsilon = 10^{-4}\left\|b-a\right\|$.
 После выполнения заданного числа итераций или достижения методом глобального поиска заданной точности решение уточнялось Hook-Jeeves local method \cite{HookJeeves} с точностью $\epsilon = 10^{-4}\left\|b-a\right\|$.
+
+Детальное исследование области значений параметра регуляризации в окрестности $\alpha$ равного нулю (от 0.1 до 0.01 с шагом 0.01 и от 0.01 до 0.001 с шагом 0.001) выявило убывание значения минимума с уменьшением параметра $\alpha$.
+Лучшее решение $0.540750$ было получено при параметре регуляризации $\alpha = 0.001$.
 
@@
 
-Детальное исследование области значений параметра регуляризации в окрестности alpha равного нулю (от 0.1 до 0.01 с шагом 0.01 и от 0.01 до 0.001 с шагом 0.001) выявило убывание значения минимума с уменьшением параметра $alpha$, см.
-рис.
-...
-.
-
 \section{Conclusions and Future Work}

### 2023-01-27 2023/PAVT_Ufa/PAVT2023.tex
fix architecture claster

@@
 Численные эксперименты проводились при помощи параллельного асинхронного алгоритма, изложенного в Section 4, с использованием $160$ процессов ($159$ процессов вычисляли значения целевой функции, $1$ процесс управлял алгоритмом).
-Для проведения экспериментов использовался UNN supercomputer Lobachevsky (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node).
+Для проведения экспериментов использовался UNN supercomputer Lobachevsky (CentOS 7 Intel(R) Xeon(R) Silver 4310T CPU @ 2.30GHz RAM 64Gb).
 Асинхронный алгоритм глобальной оптимизации был реализован на C++ (GCC 9.5.0 and Intel MPI were used); вычисление значений целевой функции было реализовано в Python 3.9.

### 2023-01-27 2023/PAVT/ParallelDecisionTree.tex
before translation of the last section

@@
 
-% переводить до этого раздела
+% перевести этот раздел
 \section{Результаты экспериментов}\label{SecR}
 
-Вычислительные эксперименты были проведены на Lobachevsky supercomputer. Один узел кластера состоит из двух процессоров Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM.
+Вычислительные эксперименты были проведены на Lobachevsky supercomputer. На узлах суперкомпьютера установлены по два процессора Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM.
 
@@
 
-%Ниже приведены результаты сравнения двух параллельных алгоритмов --- асинхронного GSA и его модификации с использованием decision tree и локального метода Хука-Дживса.
-Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{fio_bib19}.
-Критерием остановки служило попадание точки очередного испытания trial в $\varepsilon$ окрестность истинного глобального минимума.
-%В таблице \ref{tab:2} представлено среднее количество итераций и ускорение по времени относительно последовательного запуска.
-Распараллеливание  осуществлялось с применением  MPI.
-При этом запуск производился на 4 процессах, а следовательно, на каждой итерации вычислялось 3 точки.
-Значение в скобках (если оно есть) показывает количество не решенных задач.
-Это означает, что было достигнут максимум по числу проводимых итераций, но в окрестность глобального минимума точка так и не попала.
-
-Ниже приведены результаты сравнения двух параллельных алгоритмов --- асинхронного global search algorithm (GSA) и его модификации с использованием decision tree для выявления областей притяжения локальных минимумов (Decision Tree).
+Далее приведены результаты сравнения двух параллельных алгоритмов --- асинхронного global search algorithm (GSA) и его модификации с использованием decision tree для выявления областей притяжения локальных минимумов (Decision Tree).
 Численное сравнение проводилось на классах функций GKLS двух видов (Simple and Hard) размерности 2, 3, 4 и 5 из \cite{fio_bib19}.
@@
 
-Как видно из таблицы ускорение по числу итераций достаточно велико, причем на задачах любой размерности.
-
-%\section{Заключение}\label{SecC}
-
-%Методы глобальной оптимизации являются областью активных научных исследований
+Как видно из таблицы iteration speedup достаточно велико, причем на задачах любой размерности.
 

### 2023-01-27 2023/PAVT/ParallelDecisionTree.tex
some corrections in English

@@
 \institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
-\email{konstantin.barkalov@itmm.unn.ru, ilya.lebedev@itmm.unn.ru, rotor12587@mail.ru}
+\email{ \{konstantin.barkalov,ilya.lebedev,silenko\}@itmm.unn.ru}
 }
@@
 In the present work, the solving of the multidimensional global optimization problems using decision tree to reveal the attractor regions of the local minima is considered.
-The objective function of the  problem is defined as a 'black box''.
+The objective function of the  problem is defined as a ``black box''.
 We assume it only to satisfy Lipschitz condition with unknown constant.
 Global search algorithm is applied for the search of global minimum in the problems of such type.
-In the present work, we propose a method for selecting the nearness of local extrema of the objective  function based on analysis of accumulated search information.
-Conducting such an analysis using machine learning tools allows making a decision to run a local  method, which can speed up the convergence of the algorithm.
+In the present work, we propose a method for selecting the vicinity of local extrema of the objective  function based on analysis of accumulated search information.
+Conducting such an analysis using machine learning techniques allows making a decision to run a local  method, which can speed up the convergence of the algorithm.
 This suggestion was confirmed by the results of numerical experiments demonstrating the speedup  when solving a series of test problems.
@@
 % the abstract section using the \keywords{...} command.
-\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$  Decision tree}
+\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Machine learning $\cdot$ Decision tree}
 \end{abstract}
@@
 In the present work, we consider the parallel algorithms for solving the multidimensional global optimization problems.
-Such problems arise often in the cases, when it is necessary to select the values  of parameters of the mathematical model being investigated, which the results of modeling fit best to  the experimental data at.
-For solving the problems of the class specified, many algorithms are known: from the ones based on the  idea of random search \cite{fio_bib1, fio_bib2, fio_bib3} metaheuristic algorithms to the deterministic  algorithm guarantying convergence to the global minimum \cite{fio_bib4, fio_bib5, fio_bib6}.
+Such problems often arise in the cases, when it is necessary to select the values  of parameters of the mathematical model being investigated, in which the simulation results best fit the experimental data.
+For solving the problems of this class, many algorithms are known: from the metaheuristic algorithms, based on the  idea of random search \cite{fio_bib1,fio_bib2,fio_bib3}, to the deterministic algorithms guarantying convergence to the global minimum \cite{fio_bib4,fio_bib5,fio_bib6}.
 
-Since in real global optimization problems each computing of the function value (hereafter called  \textit{a trial}) is a very computation-costly operation, one has to reduce the number of such trials.
-It  can be achieved by intentional choice of variants in the course of search for the optimal solution cutting  off unpromising search subdomains and investigating only these ones, which the solution of the  problem can be found in.
+Since in real global optimization problems each computing of the function value (hereafter called  \textit{a trial}) is a computation-costly operation, one has to reduce the number of such trials.
+It  can be achieved by intentional choice of variants in the course of search for the optimal solution cutting  off unpromising search subdomains and investigating only these ones, in which the solution of the  problem can be found.
 The global search algorithm (GSA) is based on this idea \cite{fio_bib7}.
-In  the present work, we tried to combine GSA and a local optimization method (Hooke-Jeeves method)  to reduce the number of trials executed.
+In  the present work, we tried to combine GSA and a local optimization method (Hooke-Jeeves pattern search method) to reduce the number of trials executed.
 The decision on the run of the local method will be made using a decision tree.
 
-The local optimization algorithms are intended for determining only one of local extrema within a set  of feasible solutions, which the objective function takes the extremal (maximum or minimum) value in.
-Among the local methods,  the zero-order methods and the gradient methods (of the 1\textsuperscript{st} order and of the  2\textsuperscript{nd} one) are distinguished traditionally \cite{fio_bib8,fio_bib9}.
-Note that the application of the first-order methods (not to mention the second one) in the problems  with the “black-box” type functions is difficult since the problem of numerical estimating the gradient  arises here.
-Therefore, further we will consider the zero-order methods only.
+The local optimization algorithms are intended for determining only one of local extrema within a set  of feasible solutions, in which the objective function takes the extremal (maximum or minimum) value.
+Among the local methods,  the zero-order methods and the gradient methods (of the 1\textsuperscript{st} order and of the  2\textsuperscript{nd} one) are traditionally distinguished \cite{fio_bib8,fio_bib9}.
+Note that the application of the first-order methods (not to mention the second one) in the problems  with the ``black-box'' type functions is difficult since the problem of numerical estimating the gradient arises here.
+Therefore, further we will consider only the zero-order methods.
 
-In order to determine, which moment it is best to run the local method within the framework of the  global search at we will use an algorithm based on a decision tree.
+In order to determine,  at which moment it is best to run the local method within the framework of the  global search, we will use an algorithm based on a decision tree.
 At the first search stage, the search information is accumulated in the form of the results of the trials  executed.
-We used the trial results for training the decision tree that allows obtaining a piecewise linear  approximation of the objective function and predicting the objective function behavior on its base.
-To do so, we will compare the function values at the points neighboring to the one, which we consider  to be suspicious for a local minimum and, based on the piecewise linear approximation, make a decision  whether the next trial point falls into the attractor area of a local minimum  or not.
+We used the trial results for training the decision tree that allows obtaining a piecewise constant approximation of the objective function and predicting the objective function behavior on its base.
+To do so, we will compare the function values at the points neighboring to the one, which we consider  to be suspicious for a local minimum, and make a decision  whether the next trial point falls into the attractor region of a local minimum or not.
 
@@
 In the next section,  the features of the asynchronous parallelization scheme of the new algorithm are discussed.
-Section \ref{SecR} contains the results of experiments , carried out using a parallel computer system  with distributed memory.
-In Conclusion, main conclusions on the work done are presented.
+Section \ref{SecR} contains the results of experiments, carried out using a parallel computer system  with distributed memory, and a brief conclusion.
 
@@
 Let us consider a problem of searching the global minimum of a function $\varphi(y)$ in a hyperinterval  $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1\le\ i\le\ n \}$. 
-
 \begin{equation} \label{sec:problem}   
@@
 \end{equation}
-
 where $a,b \in R$ are given vectors.
 
-Also, assume the function to satisfy the Lipschitz condition with \textit{a priori} unknown constant  $L$ that corresponds to limited variation of the function values at limited variation of the argument.
-generating the changes in the system being simulated.
-
+Also, assume the function to satisfy the Lipschitz condition with \textit{a priori} unknown constant $L$, that corresponds to limited variation of the function values at limited variation of the argument.
 \begin{displaymath} 
@@
 \end{displaymath}
+This suggestion can be interpreted (with respect to the applied problems ) as the reflection of limited  power generating the changes in the system being simulated.
 
-This suggestion can be interpreted (with respect to the applied problems ) as the reflection of limited  power
-
-The numerical solving of the problem (\ref{sec:problem})  is reduced to constructing an estimate  $ y_k^\ast\in\ D$ matching some understanding of nearness to the point $y^\ast$ (for example,  ${||y^\ast-y}_k^\ast||\le\ \varepsilon$ where $\varepsilon\geq0$ is a predefined precision) based on a  finite number $k$ of the objective function values computed.
-With respect to the class of the problems  considered, it is suggested that the optimized function $\varphi(y)$ can be defined algorithmically, as  a result of executing some subroutine or library.
+The numerical solving of the problem (\ref{sec:problem})  is reduced to constructing an estimate  $ y_k^\ast\in D$ matching some concept of nearness to the point $y^\ast$ (for example,  ${||y^\ast-y}_k^\ast||\le\ \varepsilon$, where $\varepsilon\geq0$ is a predefined accuracy) based on a  finite number $k$ of the objective function values computed.
+With respect to the class of the considered problems, it is suggested that the objective function $\varphi(y)$ can be defined algorithmically, as a result of executing some subroutine or library.
 
@@
 The decision rules of the grid building algorithm are constructed in such a way that the next  grid point corresponds to the global minimum point of the mathematical expectation of the function  values.
-This point is stored in the list of known values, and the iterations are repeated until one of  selected stop criteria is satisfied: either the distance between the trial points becomes less than a  predefined value or a preset maximum number of iterations is achieved \cite{fio_bib10}.
+This point is stored in the list of known values, and the iterations are repeated until the stop criteria is satisfied: either the distance between the trial points becomes less than a  predefined value or a preset maximum number of iterations is achieved \cite{fio_bib10}.
 
@@
 These  ones allow reducing a multidimensional optimization problem in the domain $D$  to a one-dimensional  minimization problem within the interval $[0, 1]$
-
 \begin{displaymath}
-	\varphi(y(x^\ast))\ =\ min\{\varphi(y(x)):\ x\epsilon[0,\ 1]\}
+	\varphi(y(x^\ast))\ =\min\{\varphi(y(x)): x\in [0,1]\},
 \end{displaymath}
-
 where the function $\varphi(y(x^\ast))$ satisfies more general H{\"o}lder condition
-
 \begin{displaymath}
@@
 
-Therefore, instead of the initial problem of minimizing the function $\varphi(y)$ in the domain $D$,  one can consider the minimization of the one-dimensional function $f(x)\ =\ \varphi(y(x))$ satisfying  the H{\"o}lder condition for $ x\in [0,1]$.
+Therefore, instead of the initial problem of minimizing the function $\varphi(y)$ in the domain $D$,  we can consider the minimization of the one-dimensional function $f(x)=\varphi(y(x))$ satisfying  the H{\"o}lder condition for $ x\in [0,1]$.
 
@@
 
-If $n \geq 1$ iterations are completed, which $k=k(n)$ executed trials in the points $x^i, 1\leq i\leq  k$ correspond to, the points $x^{k+1},\ldots,x^{k+p}$ of the search trials at the next $(n+1)^{\rm  th}$ iteration will be computed as the result of performing the following operations.
+If $n \geq 1$ iterations are completed, which correspond to $k=k(n)$ trials executed at the points $x^i, 1\leq i\leq  k$, then the points $x^{k+1},\ldots,x^{k+p}$ of the search trials at the next $(n+1)^{\rm  th}$ iteration will be computed as the result of performing the following operations.
 
@@
 
-	\item  Renumber (by the lower indices) the points $x^i, 1\leq i\leq k$ as well as the boundary points of  the interval [0,1] in the order of increasing coordinate:  
+	\item  Renumber (by the lower indices) the points $x^i, 1\leq i\leq k$ as well as the boundary points of  the interval [0,1] in increasing order of the coordinate  
 	\begin{equation} 
@@
 	\end{equation} 
-
 	where $r>1$ is a parameter of algorithm.
-This parameter controls the reliability of the algorithm: the  higher values of $r$ ensure guaranteed finding of the global minimum, the choice of lower value ---  speeds up the convergence of the algorithm.
+This parameter controls the reliability of the algorithm: the  higher values of $r$ ensure guaranteed finding of the global minimum, the choice of lower value speeds up the convergence of the algorithm.
 
 	\item  For each interval $(x_{i-1},x_i), 1\leq i\leq k+1,$ compute the value $R(i)$ called \textit{a  characteristic} of the interval according to the formulae 
-
 	\begin{equation} 
@@
 	\end{equation} 
-
 	\begin{equation} 
@@
 	\end{equation} 
-
 	where \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
 
-	\item   Arrange the characteristics $R\left(i\right),\ 1\leq i \leq k+1,$ in the order of non-increasing
-
+	\item   Arrange the characteristics $R\left(i\right),\ 1\leq i \leq k+1,$ in the non-increasing order 
 	\begin{equation} 
@@
 	\end{equation} 	
-
-	and select $p$ intervals with the indices $t_j,\ 1\le\ j\le\ p$ with the highest values of characteristics.
+	and select $p$ intervals with the indices $t_j,\ 1\le\ j\le\ p$ with the largest values of characteristics.
 
 	\item Compute the points $x^{k+j},\ 1\leq j\leq p$ in the selected intervals according to the formulae
-
 \begin{equation}
@@
 
-The next $p$ trials are executed in parallel in the points $x^{k+j},\ 1\leq j\leq p$ computed according  to the formulae (\ref{agp5_x1}), (\ref{agp4_xi}).
+The next $p$ trials are executed in parallel at the points $x^{k+j},\ 1\leq j\leq p$, computed according  to the formulae (\ref{agp5_x1}), (\ref{agp4_xi}).
 Upon completing the trials, the results of these ones  are stored in the information database, and the algorithm goes to computing new trial points.
-
 Note that as a rule the process of executing a trial in the applied optimization problems is much more  computation-costly as compared to computing the trial point.
@@
 The algorithm stops in the case if the condition \(\Delta_{t_j} < \varepsilon\) is satisfied for even a  single value $t_j,\ 1\le\ j\le\ p$ from (\ref{agp4_R_sort}).
-This stop criterion (along with the criterion limiting the number of executed iterations usual for the  iteration methods) is used in the applied optimization problems, which the global minimum point  $x^*$ is unknown \textit{a priori} in.
+This stop criterion (along with the criterion limiting the number of executed iterations usual for the  iteration methods) is used in the applied optimization problems, in which the global minimum point $x^*$ is unknown a priori.
 
-When solving the test problems, which the global minimum point $x^*$ is known in one can use the  stop criterion upon falling into the nearness of the global minimum as well.
-In this case, the method  stops if the condition $\left|x_{t_j}-\ x^\ast\right| < \varepsilon$ is satisfied even for one value of $t_j,\  1\le\ j\le\ p$ from (\ref{agp4_R_sort}).
+When solving the test problems, in which the global minimum point $y^*$ is known, we can use the stop criterion upon falling into the vicinity of the global minimum as well.
+In this case, the method  stops if the condition $\left\|y(x_{t_j})-\ y^\ast\right\| < \varepsilon$ is satisfied even for one value of $t_j,\  1\le\ j\le\ p$, from (\ref{agp4_R_sort}).
 
-As the final estimate of the globally optimized solution of the considered problem, the values
-
+As the final estimate of the global optimizer of the considered problem, the values 
 \begin{equation} 
@@
 \end{equation} 
-
 are taken.
@@
 The substantiation of this method of organization of computations see in \cite{fio_bib20}.
-The  modifications taking into account the presence of  constraints (inequalities) in the problem as well as  the information on the derivative of the objective function  are presented in \cite{fio_bib12, fio_bib9,  fio_bib11}.
+The  modifications taking into account the presence of inequality constraints as well as the information on the derivative of the objective function are presented in \cite{fio_bib12, fio_bib9,  fio_bib11}.
 
@@
 
-The Hooke--Jeeves method belongs to the class of zero-order methods.
+The Hooke--Jeeves pattern search algorithm belongs to the class of zero-order methods.
 Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{fio_bib14, fio_bib15}.
 
-The investigating search is performed as follows:  \begin{enumerate} \item	The step value is determined (it is different for each coordinate and can be varied in the course  of search).
-\item	The search step is considered to be successful if the value of the objective function in the check  point doesn't exceed the value of the objective function in the initial point.
+The investigating search is performed as follows. 
+\begin{enumerate} 
+\item	The step value is determined (it is different for each coordinate and can be varied in the course  of search).  
+\item	The search step is considered to be successful if the value of the objective function at the check  point doesn't exceed the value of the objective function at the initial point.  
 \item	Otherwise, it is necessary to return to the previous item and make a step in the reverse direction.  
-\item	After making the steps in all $N$ coordinates, the investigating search is completed.
-The point  obtained is called the base point.
+\item	After making the steps in all $N$ coordinates, the investigating search is completed. The point  obtained is called the base point. 
 \end{enumerate}
@@
 The decision tree is a tool used for an automated analysis of big data arrays, which is applied in the  machine learning.
-Decision tree is a binary tree (a tree, which each non-leaf node has two child nodes  in).
+Decision tree is a binary tree (a tree, in which each non-leaf node has two child nodes).
 It can be used for solving the classification problems as well as the regression ones.
@@
 Therefore, the approximating function obtained is a piecewise-constant one.
-In our case, the decision tree constructs the function $\varphi(y)$ as a piecewise-constant  approximation $\varphi(y)$  in multidimensional space.
-Let us denote the value  computed by the  decision tree in the point $y$ as $z' = \psi(y)$.
+In our case, the decision tree constructs the function $\psi(y)$ as a piecewise-constant  approximation of $\varphi(y)$ in the search domain.
+Let us denote the value  computed by the  decision tree at the point $y$ as $z' = \psi(y)$.
 
-When implementing the algorithm for solving the multidimensional global optimization problems using  decision tree, we used the algorithms from OpenCV library to find the attractor areas of the local  minima.
-OpenCV is a library of the algorithms of computer vision, image processing, and  general-purpose numerical algorithms with open code.
+When implementing the algorithm for finding the attraction areas of the local  minima, we used the algorithms from OpenCV library to construct decision tree.
+OpenCV is an open source library of the algorithms of computer vision, image processing, and  general-purpose numerical algorithms.
 More details on the decision tree can be found  in \cite{fio_bib16}.
 
-\section{Combination of local optimization method and GSA for solving multidimensional  problems}\label{SecGSAL}
+\section{Combination of local and global search algorithms for solving multidimensional  problems}\label{SecGSAL}
 
-In the current section, let us present a detailed description how we used the decision tree for finding  the attractor areas of the local extrema.
+In the current section, let us present a detailed description how we used the decision tree for finding  the attraction areas of the local extrema.
 In the course of running GSA, it is necessary to determine whether it is worth to use current point as a  start one for the local method or not.
 To do so, one can check the points of adjacent trials.
-If among  these ones there are no points such that the function values in which are lower than in current one, one  can suggest that we are in the attractor of a local minimum.
-In this case, one can run the local method  from current point, which will converge to a local minimum of the function rapidly.
-One can do it in the multidimensional space only, not in one-dimensional interval after the  dimensionality reduction.
+If among  these ones there are no points at which the function values are lower than in current one, one  can suggest that we are in the attractor of a local minimum.
+In this case, we can run the local method  from current point, which will rapidly converge to a local minimum of the function.
+We can do it in the multidimensional space only, not in one-dimensional interval after the  dimensionality reduction.
 First, the reduced function $\varphi(y(x))$ changes its properties: one local minimum in the  multidimensional space may divide into a set of minima after the dimensionality reduction.
 Second, after the use of the mapping $y(x)$, we can lose the information on the mutual arrangement of  the points in the original space.
-The points located close to each other in the multidimensional space  may appear to be separated essentially in the one-dimensional interval.
+The points located close to each other in the multidimensional space  may appear to be essentially separated in the one-dimensional interval.
 
@@
 
-Therefore, we will determine the attractor areas of the local extrema in the original multidimensional  space.
-However, the determining of the adjacent points in the multidimensional space is  computational-costly enough.
+Therefore, we will determine the attraction areas of the local extrema in the original multidimensional space.
+However, the determining of the adjacent points in the multidimensional space is quite expensive in terms of computing resources.
 To do so, we will use the decision tree.
 
-After completing a definite number of trials (for example, $100\ \ast\ N$), let us use all accumulated  points to initialize a suitable data structure and training the decision tree on its base.
+After completing a certain number of trials (for example, $100\ \ast\ N$), let us use all accumulated  points to initialize a suitable data structure and training the decision tree on its base.
 To make the  determining the adjacent points easier, we construct a uniform grid with certain step in each coordinate.
-Afterwards, we compute the values of the approximation of the objective function in these points  using the decision tree.
+Next, we compute the values of the approximation of the objective function in these points  using the decision tree.
 Now, having a uniform grid of points and knowing the function values in each  point, let us find the closest point to the initial one (from the viewpoint of the Euclidian distance).
-This  closest point is a projection of the initial point onto the uniform grid, which its neighbors can be  determined easily on.
-Since the decision tree constructs the approximation of the initial function (even  piecewise constant), we can estimate the function values in the areas, which the trials haven’t been  executed earlier in.
+This  closest point is a projection of the initial point onto the uniform grid, on which its neighbors can be  determined easily.
+Since the decision tree constructs the approximation of the initial function (even  piecewise constant), we can estimate the function values in the areas, in which the trials haven't been  executed earlier.
 
@@
 If even a single  neighbor has a lower function value than current point, there is no need to run the local method.
-If in  even a single neighboring point the function value is the same as in the current point, it is also necessary  to check (i.
-e.
-to find all its neighbors and to heck these ones in the same way).
-Only in the case when  all function values in the neighboring points are greater than in the current point, we can run the local  method.
+If at  even a single neighboring point the function value is the same as at the current one, this point must be also checked (i.e.
+to find all its neighbors and to check these ones in the same way).
+Only in the case when  all function values at the neighboring points are greater than at the current point, we can run the local  method.
 
@@
 
-The method of parallel computing organization described in  previous section implies synchronous  performing of $p$ search trials in the points computed at current iteration.
-The next iteration will be  started upon completing all trials only.
-In the case of different computation costs of executing the trials in different points of the search  domain, this approach may lead to a misbalance of the computing jobs.
+The method of parallel computing described in  previous section implies synchronous  performing of $p$ search trials at the points computed at current iteration.
+The next iteration will be  started only upon completing all trials.
+In the case of different computation costs of executing the trials at different points of the search  domain, this approach may lead to a disbalance of the computing jobs.
 This drawback can be corrected by introduction of asynchrony.
-The idea of the asynchronous scheme is not to wait for completing all $p$ trials but to initiate the  execution of a new trial immediately upon completing at one trial.
+The idea of the asynchronous scheme is not to wait for completing all $p$ trials but to initiate the  execution of a new trial immediately upon completing one of the trials.
 
-Let us assume that the master process computes one point of the next trial at every iteration and sends  it to the performing process to execute the trial.
-Note that executing a trial by the performing process in  the applied optimization problems is much more computation costly than the choice of a new trial point  by the master process.
-Therefore, the idle time of the performing processes will be negligible.
-In this  case (unlike the synchronous parallel algorithms) total number of trials executed by each performing  process will depend on the computation costs for executing particular trials and cannot be estimated in  advance.
-When describing the parallel algorithm, let us assume that we have $p+1$ computational  processes at our disposal: one master process and $p$ performing ones.
+Let us assume that the master process computes one point of the next trial at every iteration and sends  it to the worker process to execute the trial.
+Note that executing a trial by the worker process in  the applied optimization problems is much more computation costly than the choice of a new trial point  by the master process.
+Therefore, the idle time of the processes will be negligible.
+In this  case (unlike the synchronous parallel algorithms) total number of trials executed by each worker  process will depend on the computation costs for executing particular trials and cannot be estimated in  advance.
+When describing the parallel algorithm, let us assume that we have $p+1$ computational  processes at our disposal: one master and $p$ workers.
 
 In the beginning of search, the master process (assume it to have the identifier $0$) initiates parallel  execution of $p$ trials in $p$ different points of the search domain.
+Two of these points are the boundary ones, the rest are the internal points, i.e.
+at the points  $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$ where $x^1=0,x^p=1,x^i\in\left(0,1\right),i=2,\ldots,p-1$.
 
-Two of these points are the boundary ones, the rest are the internal points i.
-e.
-in the points  $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$ where $x^1=,x^p=,x^i\in\left(0,1\right),i=2,\ldots,p$.
+Now assume $k\geq0$ trials to be completed, and the worker processes are performing the trials at the points $y\left(x^{k+1}\right),\;y\left(x^{k+2}\right),\ldots,\;y\left(x^{k+p}\right)$.
 
-Now assume $k\geq0$ trials to be completed, and the performing processes are performing the trials in  the points $y\left(x^{k+1}\right),\;y\left(x^{k+2}\right),\ldots,\;y\left(x^{k+p}\right)$.
+Each worker process having completed the computing of the function at some time moment sends  the computed value to the master process.
+In turn, the master process selects a new point  $x^{k+p+1}$ for the worker process according to the rules described below.
+Note that in this case  we will have a set of preimages of the trial points  $I_k=\left\{x^{k+1},x^{k+2},\ldots,x^{k+p}\right\}$, in which the trials have been already started but haven't been completed yet.
 
-Each performing process having completed the computing of the function at some time moment sends  the computed value to the master process.
-In turn, the master process selects a new point  $x^{k+p+1}$ for the performing process according to the rules described below.
-Note that in this case  we will have a set of preimages of the trial points  $I_k=\left\{x^{k+1},x^{k+2},\ldots,x^{k+p}\right\}$ which the trials have been started already but haven’t been completed yet in.
+So far, the parallel asynchronous global optimization algorithm using decision tree to find the local  extrema attractors consists of the following steps.
 
-So far, the parallel asynchronous global optimization algorithm using decision tree to find the local  extrema attractors consists of the following steps:
-
-\begin{enumerate} \item Arrange n the increasing order (by the lower indices) the set of preimages of the trial points
-
+\begin{enumerate} \item Arrange in the increasing order (by the lower indices) the set of preimages of the trial points
 	\begin{displaymath} 
@@
 	\end{displaymath} 
-
-	containing all points, which the trials either have been completed or are being executed in i. e. get an  ordered set
-
+	containing all points, in which the trials either have been completed or are being executed, i.e. get an  ordered set 
 	\begin{displaymath} 
@@
 	\item Compute the values 
-
 	\begin{displaymath} 
@@
 	\end{displaymath} 
-
 	\begin{displaymath} 
@@
 	\end{displaymath} 
-
 	where $ z_i=\varphi\left(y\left(x_i\right)\right)$ if $x_i\notin I_k,\;1\le i\le k+p$.
-The values $z_i$ in  the points $x_i\in I_k$ are undefined since the trials in the points $x_i\in I_k$ haven’t been completed  yet.
+The values $z_i$ at the points $x_i\in I_k$ are undefined since the trials at the points $x_i\in I_k$ haven't been completed yet.
 If the value of $M$ equals to $0$, set $M=1$.
 
-	\item Juxtapose each interval $\left(x_{i-1},x_i\right),\;x_{i-1}\notin I_k,x_i\notin I_k,\;2\le i\le  k+p$ with its characteristic $R\left(i\right)$ computed according to the formula
-
+	\item Juxtapose each interval $\left(x_{i-1},x_i\right),\;x_{i-1}\notin I_k,x_i\notin I_k,\;2\le i\le  k+p$, with its characteristic $R\left(i\right)$ computed according to the formula 
 	\begin{displaymath} 
@@
 	\end{displaymath} 
-
 	where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$ and $ r>1$ is the reliability parameter of the method.
 
-	\item Select the interval $\left[x_{t-1},x_t\right]$, which the highest characteristics correspond to, i. e.
-
+	\item Select the interval $\left[x_{t-1},x_t\right]$, which corresponds to the highest characteristic, i.e.
 	\begin{displaymath} 
@@
 
-	\item Define a new trial point $y^{k+p+1}=y\left(x^{k+p+1}\right)$, the preimage of which is  $x^{k+p+1}\in\left(x_{t-1},x_t\right)$ according to the formula
-
+	\item Define a new trial point $y^{k+p+1}=y\left(x^{k+p+1}\right)$, the preimage of which is  $x^{k+p+1}\in\left(x_{t-1},x_t\right)$, according to the formula 
 \begin{displaymath}
@@
 
-	\item If $k\ <\ 100\ast\ N$, return to Step 1.
+	\item If $k\ <\ 100 N$, return to Step 1.
 
@@
 
-Upon computing the next trial point, the master process adds it to the set $I_k$ and sends to the  performing process, which initiates the new trial in this point.
+Upon computing the next trial point, the master process adds it to the set $I_k$ and sends to the  worker process, which initiates the new trial in this point.
 The master process stops the algorithm, if one of the two conditions is satisfied: $\Delta_t<$ or  $k+p>K_{max}$.
-The real value $\epsilon>0$ and the integer value $K_{max}>0$ are the parameters of the algorithm  and correspond to the precision of the search for the solution and the maximum number of trials,  respectively.
+The real value $\epsilon>0$ and the integer value $K_{max}>0$ are the parameters of the algorithm  and correspond to the accuracy of the search and the maximum number of trials, respectively.
 

### 2023-01-29 2023/PAVT_Ufa/PAVT2023.tex
translate to english

@@
 
-Так как математической моделью реакции является система дифференциальных уравнений, найти значения констант, входящих в нее, можно только численно (см., например, \cite{RSD}).
-Отметим, что в таких задачах целевая функция, как правило, является многоэкстремальной, т.е.
-имеет много локальных экстремумов наряду с глобальным.
+Since the mathematical model of the chemical reaction is a system of differential equations, it is only possible to find the values of the constants in this system numerically (see, e.g., \cite{RSD}).
+Note that the objective function in such problems is usually multiextremal, i.e.
+it has many local extrema along with a global one.
 %Часть ННГУ
-Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) существенно отличаются от методов локального поиска (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}).
-С алгоритмической точки зрения их можно разделить на два класса: метаэвристические и детерминированные.
-Метаэвристические алгоритмы, в основном, основаны на имитации процессов, протекающих в живой природе.
-Характерными примерами таких алгоритмов являются simulated annealing, evolution and genetic algorithms и т.д.
+Numerical methods for solving such multiextremal problems (global optimization methods) differ significantly from local search methods (see, e.g., \cite{Sergeyev2017, PaulaviciusZilinskas2014}).
+From an algorithmic point of view, they can be divided into two classes: metaheuristic and deterministic.
+Metaheuristic algorithms are mainly based on the simulation of processes occurring in living nature.
+Typical examples of such algorithms are simulated annealing, evolution and genetic algorithms, etc.
 (see, e.g., \cite{Battiti2009,Eiben2015}).
-В силу своей относительной простоты метаэвристические алгоритмы более популярны у исследователей, чем детерминированные.
-Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может располагаться далеко от глобального \cite{Kvasov2018}.
+Due to their relative simplicity, metaheuristic algorithms are more popular with researchers than deterministic ones.
+However, the solution to a problem found by a metaheuristic algorithm is, generally speaking, local and may be well away from the global solution \cite{Kvasov2018}.
 
-Предполагая наличие некоторых дополнительных свойств целевой функции, можно построить эффективные детерминированные методы поиска глобального решения.
-Например, можно предположить, что отношение приращения функции к соответствующему приращению аргумента не может превышать некоторого порога.
-В таком случае функции называются липшицевыми, а сама задача -- задачей липшицевой глобальной оптимизации .
-Данная статья продолжает цикл работ, в которых проводится исследование и модификация параллельных методов липшицевой оптимизации, предложенных в \cite{Strongin2000}, при их использовании для решения обратных задач химической кинетики.
+Assuming some additional properties of the objective function, it is possible to construct efficient deterministic methods for finding a global solution.
+For example, we can assume that the ratio of the increment of the function to the corresponding increment of the argument cannot exceed some threshold.
+In this case, the functions are called Lipschitz functions, and the problem itself is a Lipschitz global optimization problem.
+This paper continues the series of papers where the parallel methods of Lipschitz optimization proposed in \cite{Strongin2000} are investigated and modified in their application to solving inverse problems of chemical kinetics.
 
-Основная часть статьи имеет следующую структуру.
-Описание математической модели исследуемой химической реакции приведено в Section 2.
-Формальная постановка задачи липшицевой глобальной оптимизации и общая схема поисковых алгоритмов приводятся в Section 3.
-В Section 4 изложена схема предлагаемого асинхронного параллельного алгоритма для решения многоэкстремальных задач.
-Результаты численного решения the inverse problem of chemical kinetics обсуждаются в Section 5.
+The main part of the paper is organized as follows.
+Section 2 describes the mathematical model of the chemical reaction under study.
+The formal statement of the Lipschitz global optimization problem and the general scheme of search algorithms are given in Section 3.
+In Section 4, a scheme of the proposed asynchronous parallel algorithm for solving multiextremal problems is presented.
+The results of numerical solution of the inverse problem of chemical kinetics are discussed in Section 5.
 
@@
 
-Как уже говорилось выше, задача идентификации значений параметров математической модели может рассматриваться как Lipschitz global optimization problem. 
-С формальной точки зрения  данная задача является задачей математического программирования вида
+As it was mentioned above, the problem of identifying mathematical model parameter values can be considered a Lipschitz global optimization problem.
+From the formal point of view this problem is a mathematical programming problem of the form
 \begin{gather}
@@
 \end{gather}
-где $a,\; b$ есть заданные векторы, $a,b\in R^N$, $\varphi(y)$ целевая функция, которая удовлетворяет the Lipschitz condition
+where $a,\; b$ are the given vectors, and $a,b\in R^N$, $\varphi(y)$ is the objective function which satisfies the Lipschitz condition
 \begin{equation}\label{Lip}
@@
 
-Функция $\varphi(y)$ предполагается многоэкстремальной и заданной в виде ``black box'' (т.е.
-в виде некоторой subroutine, на вход которой подается вектор параметров, а на выходом является вычисленное значение функции).
-Кроме того, подразумевается, что каждое применение этой процедуры (называемое в дальнейшем search trial) является трудоемкой операцией.
-Такая постановка задачи полностью соответствует the inverse problem of chemical kinetics.
+The function $\varphi(y)$ is assumed to be multiextremal and is given as a ``black box'' (i.e.
+as some subroutine with a vector of parameters as its input and the calculated value of the function as its output).
+Moreover, it is assumed that each application of this procedure (hereinafter referred to as search trial) is a time-consuming operation.
+This formulation of the problem is fully consistent with the inverse problem of chemical kinetics.
 
-Известен ряд алгоритмов, применимых для решения задач липшицевой оптимизации.
-Среди них можно отметить non-uniform space covering method \cite{Evtushenko2013,Evtushenko2009}, diagonal and simplicial partitions methods \cite{Zilinskas2010,Paulavicius2011}.
-В данной работе мы использовали global search algorithm, proposed by Strongin \cite{Strongin2000}.
-В рамках данного подхода исходная многомерная задача (\ref{problemN}) редуцируется к задаче одномерной оптимизации при помощи Peano-Hilbert curves.
+There are several algorithms that can be applied to solve Lipschitz optimization problems.
+These include, among others, the non-uniform space covering method \cite{Evtushenko2013, Evtushenko2009}, diagonal and simplicial partitions methods \cite{Zilinskas2010, Paulavicius2011}.
+In this paper, we used the global search algorithm proposed by Strongin{Strongin2000}.
+Under this approach, the original multidimensional problem (\ref{problemN}) is reduced to a one-dimensional optimization problem using Peano-Hilbert curves.
 
-В самом деле, используя непрерывное однозначное отображение (Peano-Hilbert curve) $y(x)$ отрезка $[0,1]$ вещественной оси на гиперинтервал $D$ из (\ref{D}) можно свести многомерную задачу (\ref{problemN}) к одномерной задаче
+In fact, using a continuous one-to-one mapping (Peano-Hilbert curve) $y(x)$ of the segment $[0,1]$ of the real axis onto the hyperinterval $D$ from (\ref{D}), we can reduce the multidimensional problem (\ref{problemN}) to a one-dimensional problem
 \[
@@
 with $ H=2 L \sqrt{N+3}$.
-Вопросы численного построения различных аппроксимаций Peano-Hilbert curve рассмотрены в \cite{Strongin2000,Sergeyev2013}.
+The numerical construction of various approximations of the Peano-Hilbert curve is discussed in \cite{Strongin2000,Sergeyev2013}.
 
-Таким образом, search trial в некоторой точке $x'\in[0,1]$ будет включать в себя сначала the construction of the image $y'=y(x')$, и лишь затем вычисление значения функции $ z' = \varphi(y')$.
+Thus, the search trial at some point $x'\in[0,1]$ will involve first the construction of the image $y'=y(x')$, and only then the computation of the function value $ z' = \varphi(y')$.
 
@@
 %Характеристический алгоритм
-Примененный нами алгоритм глобального поиска принадлежит классу characteristic algorithms, что существенно упрощает процесс его распараллеливания.
-Напомним, что а numerical optimization method belongs to the class of characteristical algorithms if the algorithmic scheme of this method can be described as follows.
+The global search algorithm we used belongs to the class of characteristic algorithms, which greatly simplifies the process of its parallelization.
+Recall that a numerical optimization method belongs to the class of characteristic algorithms if the algorithmic scheme of this method can be described as follows.
 
-At the first iteration the trials are performed at the points $x^1 = 0$ and $x^2 = 1$. Then any next $(k+1)$, $k \geq 2$, iteration og global search is executed in the following way.
+At the first iteration the trials are performed at the points $x^1 = 0$ and $x^2 = 1$. Then any next $(k+1)$, $k \geq 2$, iteration of global search is executed in the following way.
 
@@
 	\item 
-Renumber points of previous trials by increasing their $x$-coordinates (the new order is marked with subscript)
+Renumber points of the previous trials by increasing their $x$-coordinates (the new order is marked with subscript)
 \[
@@
 Calculate for every interval $(x_{i-1},x_i), 1\leq i\leq k$, a value $R(i)$ called \textit{the interval characteristic}.
-In general case, $R(i)$ can depend on the points $x^i$ and the trial rezults $z^i=f(x^i), 1 \leq i \leq k$.
+In the general case, $R(i)$ can depend on the points $x^i$ and the trial results $z^i=f(x^i), 1 \leq i \leq k$.
 
@@
 $(x_{i-1},x_i)$.
-To construct a concrete form of the interval characteristic we can use a lower envelope (or minorant) of the function to be minimized or a mathematical expectation of function values, etc. 
+To construct a concrete form of the interval characteristic, we can use a lower envelope (or minorant) of the function to be minimized or a mathematical expectation of function values, etc. 
 Most well-known global optimization methods can be formulated in accordance with this characteristical scheme, e.g.,
@@
 
-The Strongin algorithm has been constructed in the framework of the information approach to global optimization (see Strongin (1978)).
+The Strongin algorithm has been constructed in the framework of the information approach to global optimization (see \cite{Strongin1978})).
 This method has an adaptive scheme to evaluate a numerical estimate of the unknown Lipschitz constant.
@@
 
-Принадлежность Global Search Algorithm (GSA) к классу caracteristic algorithms подсказывает возможный путь его распараллеливания. 
-Как уже было отмечено, характеристика интервала $R(i)$ может рассматриваться как некоторая мера нахождения (локализации?) в нем точки глобального минимума. 
-Тогда вместо одного лучшего интервала можно выбрать сразу несколько интервалов с наибольшими характеристиками и провести в них очередные испытания параллельно.
-Более того, схема характеристических алгоритмов допускает и асинхронное распараллеливание, которое минимизирует простой процессоров в случае, если трудоемкость проведения испытаний зависит от конкретной точки области поиска.
+As the Global Search Algorithm (GSA) belongs to the class of characteristic algorithms, this suggests a possible way to parallelize it. 
+As previously mentioned, the characteristic of the interval $R(i)$ can be regarded as some measure of finding the global minimum point in this interval.
+Then, instead of a single best interval, several intervals with the highest characteristics can be chosen at once and successive trials can be carried out in these intervals in parallel.
+Moreover, the scheme of characteristic algorithms also allows asynchronous parallelization, which minimizes downtime of processors when trial complexity depends on a particular point in the search domain.
 
-Рассмотрим более детально асинхронный алгоритм глобального поиска.
-В нем реализована схема распараллеливания вида ``master/worker''.
-В процессе-мастере проводится накопление поисковой информации, оценка на ее основе константы Липшица для целевой функции, определение точек новых испытаний и распределение их по процессам-рабочим.
-Процессы-рабочие получают от мастера точки, проводят в них новые испытания и отсылают мастеру их результаты.
+Now, let us examine the asynchronous global search algorithm in more detail.
+It implements a paralleling scheme of the ``master/worker'' type.
+The master process accumulates search information, evaluates on its basis the Lipschitz constant for the objective function, determines new trial points and distributes them to the worker processes.
+Worker processes receive the points from the master, perform new trials in these points and send the results to the master.
 
-При описании параллельного алгоритма будем предполагать, что в нашем распоряжении имеется $p+1$ вычислительный процесс: один мастер и $p$ рабочих.
+When describing the parallel algorithm, we assume that we have $p+1$ computational processes: one master and $p$ worker processes.
 
-В начале поиска процесс-мастер (пусть это будет процесс No 0) инициирует параллельное проведение $p$ испытаний в $p$ различных точках области поиска, две из которых являются граничными, а остальные -- внутренними, т.е.
-в точках $\{y(x^1), y(x^2), ...,y(x^p)\}$, где $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.
+At the beginning of the search, the master process (let it be process No.
+0) initiates $p$ parallel  trials at $p$ different points of the search domain, two of which are boundary points and the rest are internal points, i.e.
+at the points $\{y(x^1), y(x^2), ...,y(x^p)\}$, where $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.
 
-Предположим теперь, что выполнено $k$ испытаний (в частности, $k$ может быть равно 0), и процессы-рабочие проводят испытания в точках $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.
+Suppose now that $k$ trials have been performed (in particular, $k$ can be 0), and the worker processes perform trials at the points $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.
 
-Если процесс-рабочий завершил проведение испытания в некоторой точке (пусть это будет точка $y(x^{k+1})$, соответствующая процессу No 1), то он пересылает процессу-мастеру результаты испытания. 
-Отметим, что в данном случае мы будем иметь set of preimages of the trial points
+If a worker process completes the trial at some point (let it be the point $y(x^{k+1})$ corresponding to process No.
+1) then it sends to the master process the results of the trial.
+Note that in this case we will have a set of preimages of the trial points
 \[
@@
 \]
-в которых испытания уже начались, но еще не завершены.
+where the trials have already started but have not yet been completed.
 
-Получив от рабочего процесса результаты испытания в точке $y(x^{k+1})$ мастер, в свою очередь, выбирает для него точку нового испытания $x^{k+p+1}$ в соответствии с правилами, которые соответствуют схеме характеристического алгоритма.
+Having received the trial results at the point $y(x^{k+1})$ from the worker process, the master in turn selects a new trial point $x^{k+p+1}$ for it according to the rules that correspond to the scheme of the characteristic algorithm.
 
-1. Перенумеровать в порядке возрастания (нижним индексом) set of preimages of the trial points 
+\begin{enumerate}
+	\item 
+Renumber the set of preimages of the trial points 
 \[
@@
 \]
-которое содержит все preimages, в которых либо проведены, либо проводятся испытания, т.е.
+containing all preimages at which the trials have either been carried out or are being carried out in ascending order (marked with subscript) , i.e.
 \[
@@
 \]
-2. Вычислить значения 
+	\item
+Calculate the values
 \[
@@
 \]
-где $z_i=\varphi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$.
-Значения $z_i$ в точках $x_i \in I_k$ являются неопределенными, т.к.
-испытания в точках $x_i \in I_k$ еще не завершены.
-Если значение $M$ получилось равным 0, то положить $M=1$.
+where $z_i=\varphi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$.
+The values of $z_i$ at the points $x_i \in I_k$ are undefined, as the trials at the points $x_i \in I_k$ have not yet been completed.
+If the value of $M$ equals 0, assume $M=1$.
 
-3.
-Каждому интервалу $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, поставить в соответствие число $R(i)$, которое называется характеристикой интервала и вычисляется по формуле
+	\item
+Assign to each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, the number $R(i)$, which is called the interval characteristic and is calculated using the formula
 \[
@@
 \]
-где $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.
+where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.
 
-4. Выбрать интервал $[x_{t-1},x_t]$, которому соответствует максимальная характеристика, т.е.
+	\item
+Select the interval $[x_{t-1},x_t]$ to which the maximum characteristic corresponds, i.e.
 \[
@@
 \]
-
-5. Определить preimage $x^{k+p+1} \in (x_{t-1},x_t)$ точки нового испытания в соответствии с формулой
+	\item
+Determine the preimage $x^{k+p+1} \in (x_{t-1},x_t)$ of the new trial point according to the formula
 \[
@@
 \]
+\end{enumerate}
 
-Сразу после вычисления точки очередного испытания $y^{k+p+1} = y(x^{k+p+1})$ процесс-мастер добавляет ее в множество $I_k$ и пересылает процессу-рабочему, который инициирует проведение испытания в ней.
+Immediately after calculating the next trial point $y^{k+p+1} = y(x^{k+p+1})$, the master process adds it to the set $I_k$ and forwards it to the worker process which initiates a trial at this point.
 
-Процесс-мастер завершает работу алгоритма при выполнении одного из двух условий: $\Delta_{t}<\epsilon$, $k+p>K_{max}$.
-Первое из них соответствует остановке алгоритма по точности, второе --- по числу испытаний. 
-Вещественное число $0<\epsilon<1$ и целое число $K_{max}>0$ являются параметрами алгоритма.
+The master process completes the algorithm when one of two conditions is met: $\Delta_{t}<\epsilon$, $k+p>K_{max}$.
+). The first one corresponds to stopping the algorithm by accuracy, the second one, by the number of trials. The real number $0<\epsilon<1$ and the integer $K_{max}>0$ are the algorithm parameters.
 
@@
 
-Предварительные эксперименты позволили оценить примерное время одного вычисления целевой функции при использовании последовательного алгоритма.
-Эксперименты проводились на архитектуре Intel(R) Core(TM) i7-10750H CPU @ 2.60GHz 2.59 GHz.
-Среднее время вычисления целевой функции составило порядка $0.2$ сек., что характеризует задачу как вычислительно сложную.
-Решение задачи последовательным алгоритмом потребовало бы не менее 1 млн.
-испытаний и, соответственно, не менее 54 часов (оценочное значение).
-Поэтому для проведения расчетов использовался только параллельный алгоритм с критерием остановки по числу испытаний $K_{max}=10^6$.
+In preliminary experiments, we estimated the approximate time to calculate one objective function using the sequential algorithm.
+The computer architecture used for the experiments was Intel(R) Core(TM) i7-10750H CPU @ 2.60GHz 2.59 GHz.
+The average calculation time of the objective function was about $0.2$ sec, which characterizes the problem as being computationally complex.
+Solving the problem with a sequential algorithm would require at least 1 million trials and thus at least 54 hours (as an estimate).
+Therefore only a parallel algorithm with a stopping criterion based on the number of trials $K_{max}=10^6$ was used for the calculations.
 
-Численные эксперименты проводились при помощи параллельного асинхронного алгоритма, изложенного в Section 4, с использованием $160$ процессов ($159$ процессов вычисляли значения целевой функции, $1$ процесс управлял алгоритмом).
-Для проведения экспериментов использовался UNN supercomputer Lobachevsky (CentOS 7 Intel(R) Xeon(R) Silver 4310T CPU @ 2.30GHz RAM 64Gb).
-Асинхронный алгоритм глобальной оптимизации был реализован на C++ (GCC 9.5.0 and Intel MPI were used); вычисление значений целевой функции было реализовано в Python 3.9.
+Numerical experiments were performed using the parallel asynchronous algorithm outlined in Section 4, running $160$ processes: $159$ processes calculated the values of the objective function (worker processes) while $1$ master process controlled the algorithm.
+The Lobachevsky  supercomputer (University of Nizhni Novgorod)  (CentOS 7.2, SLURM, two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM on the node) was used for the experiments.
+The asynchronous global optimization algorithm was implemented in C++ (GCC 9.5.0 and Intel MPI were used); the computation of objective function values was implemented in Python 3.9.
 
 %Ускорение составило примерно $230$ раз.
-Малое время решения задачи при использовании параллельного алгоритма (примерно 14 минут на 1 задачу) позволило провести полноценное исследование зависимости решения от параметра регуляризации $\alpha$.
-В таблице \ref{table_1} приведены минимальные значения objective function, найденные при соответствующем значении параметра регуляризации, время решения задачи (в секундах) и time speedup по сравнению с оценкой времени решения задачи последовательным методом.
-При запусках были использованы параметр метода $r=4.0$ из (\ref{R}) и точность $\epsilon = 10^{-4}\left\|b-a\right\|$.
-После выполнения заданного числа итераций или достижения методом глобального поиска заданной точности решение уточнялось Hook-Jeeves local method \cite{HookJeeves} с точностью $\epsilon = 10^{-4}\left\|b-a\right\|$.
+The short time for solving the problem when using the parallel algorithm (approximately 14 minutes for 1 problem) allowed a thorough investigation of the dependence of the solution on the regularisation parameter $\alpha$.
+The table \ref{table_1} shows the minimum values of the objective function found for the corresponding value of the regularization parameter, the time for solving the problem (in seconds) and the time speedup in comparison to the estimated problem solution time when using the sequential method.
+The method parameter $r=4.0$ из (\ref{R}) and the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$ were used in the runs.
+After a given number of iterations or after reaching the specified accuracy with the global search method, the solution was refined by the Hooke-Jeeves local method \cite{HookJeeves} with an accuracy of $\epsilon = 10^{-4}\left\|b-a\right\|$.
 
-Детальное исследование области значений параметра регуляризации в окрестности $\alpha$ равного нулю (от 0.1 до 0.01 с шагом 0.01 и от 0.01 до 0.001 с шагом 0.001) выявило убывание значения минимума с уменьшением параметра $\alpha$.
-Лучшее решение $0.540750$ было получено при параметре регуляризации $\alpha = 0.001$.
+A detailed study of the region of the regularization parameter values in the neighborhood of $\alpha$ equal to zero (from $0.1$ to $0.01$ in steps of $0.01$ and from $0.01$ to $0.001$ in steps of $0.001$) revealed that the value of the minimum decreases with decreasing $\alpha$.
+The best solution of $0.54075$ was obtained with the regularisation parameter $\alpha = 0.001$.
 

### 2023-01-29 2023/PAVT_Ufa/PAVT2023.tex

@@
 The main part of the paper is organized as follows.
-Section 2 describes the mathematical model of the chemical reaction under study.
-The formal statement of the Lipschitz global optimization problem and the general scheme of search algorithms are given in Section 3.
-In Section 4, a scheme of the proposed asynchronous parallel algorithm for solving multiextremal problems is presented.
-The results of numerical solution of the inverse problem of chemical kinetics are discussed in Section 5.
+Section \ref{Sec_math_mod} describes the mathematical model of the chemical reaction under study.
+The formal statement of the Lipschitz global optimization problem and the general scheme of search algorithms are given in Section \ref{Sec_GO_Problem}.
+In Section \ref{Sec_PGSA}, a scheme of the proposed asynchronous parallel algorithm for solving multiextremal problems is presented.
+The results of numerical solution of the inverse problem of chemical kinetics are discussed in Section \ref{Sec_Exp}.
 
@@
 
-\subsection{Global Optimization Problem}
+\subsection{Global Optimization Problem}\label{Sec_GO_Problem}
 
@@
 \end{gather}
-where $a,\; b$ are the given vectors, and $a,b\in R^N$, $\varphi(y)$ is the objective function which satisfies the Lipschitz condition
+where $a,b\in R^N$ are the given vectors and $\varphi(y)$ is the objective function which satisfies the Lipschitz condition
 \begin{equation}\label{Lip}
@@
 These include, among others, the non-uniform space covering method \cite{Evtushenko2013, Evtushenko2009}, diagonal and simplicial partitions methods \cite{Zilinskas2010, Paulavicius2011}.
-In this paper, we used the global search algorithm proposed by Strongin{Strongin2000}.
+In this paper, we used the global search algorithm proposed by Strongin\cite{Strongin2000}.
 Under this approach, the original multidimensional problem (\ref{problemN}) is reduced to a one-dimensional optimization problem using Peano-Hilbert curves.
@@
 
-The Strongin algorithm has been constructed in the framework of the information approach to global optimization (see \cite{Strongin1978})).
+The Strongin global search algorithm has been constructed in the framework of the information approach to global optimization (see \cite{Strongin2000})).
 This method has an adaptive scheme to evaluate a numerical estimate of the unknown Lipschitz constant.
 
-\subsection{Parallel Asynchronous Global Search Algorithm}
+\subsection{Parallel Asynchronous Global Search Algorithm}\label{Sec_PGSA}
 
-As the Global Search Algorithm (GSA) belongs to the class of characteristic algorithms, this suggests a possible way to parallelize it. 
+As the global search algorithm (GSA) belongs to the class of characteristic algorithms, this suggests a possible way to parallelize it. 
 As previously mentioned, the characteristic of the interval $R(i)$ can be regarded as some measure of finding the global minimum point in this interval.
@@
 Assign to each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, the number $R(i)$, which is called the interval characteristic and is calculated using the formula
-\[
+\begin{equation} \label{R}
 R(i)=rM\Delta_i+\frac{(z_i-z_{i-1})^2}{rM\Delta_i}-2(z_i+z_{i-1}),
-\]
+\end{equation}
 where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.
@@
 
-Numerical experiments were performed using the parallel asynchronous algorithm outlined in Section 4, running $160$ processes: $159$ processes calculated the values of the objective function (worker processes) while $1$ master process controlled the algorithm.
+Numerical experiments were performed using the parallel asynchronous algorithm outlined in Section \ref{Sec_PGSA}, running $160$ processes: $159$ processes calculated the values of the objective function (worker processes) while $1$ master process controlled the algorithm.
 The Lobachevsky  supercomputer (University of Nizhni Novgorod)  (CentOS 7.2, SLURM, two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM on the node) was used for the experiments.
@@
 The table \ref{table_1} shows the minimum values of the objective function found for the corresponding value of the regularization parameter, the time for solving the problem (in seconds) and the time speedup in comparison to the estimated problem solution time when using the sequential method.
-The method parameter $r=4.0$ из (\ref{R}) and the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$ were used in the runs.
+The method parameter $r=4.0$ from (\ref{R}) and the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$ were used in the runs.
 After a given number of iterations or after reaching the specified accuracy with the global search method, the solution was refined by the Hooke-Jeeves local method \cite{HookJeeves} with an accuracy of $\epsilon = 10^{-4}\left\|b-a\right\|$.
@@
 \begin{table}
-\caption{Исследование влияния параметра регуляризации alpha на минимальное значение функции}
+\caption{Dependence of the function minimum value $\varphi^*$ on the regularization parameter $\alpha$}
 \label{table_1}
 \begin{center}
-\begin{tabular}{cccc}
+\begin{tabular}{ccc}
 \hline\noalign{\smallskip}
- Alpha      & Minimum  & Time (sec.) & Speedup \\
+ $\alpha$      & $\varphi^*$  & Time (sec.) \\
 \hline\noalign{\smallskip}
-0.1		&	0.810454	&	806.192	&	241       \\
-0.09	&	0.837868	&	828.649	&	234       \\
-0.08	&	0.808238	&	811.852	&	239       \\
-0.07	&	0.759134	&	803.626	&	242       \\
-0.06	&	0.723404	&	810.9	&	240       \\
-0.05	&	0.729415	&	804.674	&	242       \\
-0.04	&	0.736035	&	816.387	&	238       \\
-0.03	&	0.68152		&	809.185	&	240       \\
-0.02	&	0.669092	&	155.837	&	3         \\
-0.01	&	0.632944	&	825.238	&	235       \\
-0.009	&	0.63001		&	814.48	&	239       \\
-0.008	&	0.604817	&	830.448	&	234       \\
-0.007	&	0.619546	&	807.335	&	241       \\
-0.006	&	0.612614	&	834.703	&	233       \\
-0.005	&	0.615638	&	815.45	&	238       \\
-0.004	&	0.579725	&	821.421	&	237       \\
-0.003	&	0.583529	&	831.454	&	234       \\
-0.002	&	0.547907	&	850.017	&	229       \\
-0.001	&	0.54075		&	853.898	&	227       \\
+0.1		&	0.810454	&	806.192	\\
+0.09	&	0.837868	&	828.649 \\
+0.08	&	0.808238	&	811.852	\\
+0.07	&	0.759134	&	803.626 \\
+0.06	&	0.723404	&	810.9		\\
+0.05	&	0.729415	&	804.674	\\
+0.04	&	0.736035	&	816.387	\\
+0.03	&	0.68152		&	809.185	\\
+0.02	&	0.669092	&	155.837	\\
+0.01	&	0.632944	&	825.238	\\
+0.009	&	0.63001		&	814.48	\\
+0.008	&	0.604817	&	830.448	\\
+0.007	&	0.619546	&	807.335	\\
+0.006	&	0.612614	&	834.703	\\
+0.005	&	0.615638	&	815.45	\\
+0.004	&	0.579725	&	821.421	\\
+0.003	&	0.583529	&	831.454	\\
+0.002	&	0.547907	&	850.017	\\
+0.001	&	0.54075		&	853.898	\\
 \noalign{\smallskip}\hline

### 2023-01-30 2023/PAVT/ParallelDecisionTree.tex
Поправлены результаты, добавлен перевод.

@@
 
-\begin{figure}[ht!] 
-	\begin{center} 
-		\begin{minipage}[h]{0.6\linewidth} 
-			\includegraphics[width=1\linewidth]{figure/fig2.png} 
-			\caption{A function from the GKLS generator before using Peano curves} %%  подпись к рисунку 
-			\label{fig:fig2} 
-		\end{minipage} 
-	\end{center} 
-\end{figure}
-
-\begin{figure}[ht!] 
-	\begin{center} 
-		\begin{minipage}[h]{0.6\linewidth} 
-			\includegraphics[width=1\linewidth]{figure/fig2.png} 
-			\caption{A figure with a two-dimensional decision tree} %% подпись к  рисунку 
-			\label{fig:fig2_2} 
-		\end{minipage} 
-	\end{center} 
-\end{figure}
-
 Therefore, we will determine the attraction areas of the local extrema in the original multidimensional space.
@@
 
-Fig. \ref{fig:fig2} presents a contour plot of an objective function and the trial points; Fig.  \ref{fig:fig2_2} presents corresponding decision tree.
+%Fig. \ref{fig:fig2} presents a contour plot of an objective function and the trial points; Fig.  \ref{fig:fig2_2} presents corresponding decision tree.
 
@@
 
-% перевести этот раздел
+% переводить до этого раздела
 \section{Результаты экспериментов}\label{SecR}
 
-Вычислительные эксперименты были проведены на Lobachevsky supercomputer. На узлах суперкомпьютера установлены по два процессора Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM.
+The numerical experiments were carried out on Lobachevsky supercomputer. Each supercomputer  node had two processors Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM.
 
-В экспериментах мы использовали генератор тестовых задач GKLS, который может порождать многоэкстремальные задачи оптимизации с известными свойствами: точкой глобального минимума, количеством локальных минимумов, и т.п.
+In the experiments, we used GKLS generator of test problems, which can generate multiextremal  optimization problems with known properties: the global minimum point, the number of local minima,  etc.
 
-Для демонстрации эффективности исследуемого алгоритма глобального поиска приведем результаты его сравнения с известным методами DIRECT и DIRECTl
-В таблице \ref{tab:1} приведено среднее количество итераций, выполненных соответствующими методами при решении серии задач GKLS.
-Знак «>» указывает на ситуацию, когда не все задачи были решены, здесь в скобках указано количество нерешенных задач.
-Как видно, GSA превосходит методы DIRECT и DIRECTl по среднему числу итераций, требующихся для решения задач с одинаковой точностью.
+To demonstrate the efficiency of  investigated global search algorithm, here we present the results of  comparing this one with well known methods DIRECT and DIRECTl.
+Table \ref{tab:1} presents the averaged numbers of iterations executed by respective methods when  solving a series of GKLS problems.
+The sign «>» points to the situation when not all problems are  solved.
+In these cases, the numbers of non-solved problems are given in brackets.
+As one can see,  GSA overcomes DIRECT and DIRECTl methods in average number of iterations necessary for  solving the problems with the same precision.
 
 \begin{table}[!ht]
-    \caption{Среднее число итераций}
+    \caption{Averaged number of iterations}
     \label{tab:1}
@@
     \hline
-        N & Problem class & DIRECT & DIRECTl & АГП  \\ \hline
+        N & Problem class & DIRECT & DIRECTl & GSA  \\ \hline
         4 & Simple & >47282 (4) & 18983 & 11953  \\ \hline
@@
 
-Далее приведены результаты сравнения двух параллельных алгоритмов --- асинхронного global search algorithm (GSA) и его модификации с использованием decision tree для выявления областей притяжения локальных минимумов (Decision Tree).
-Численное сравнение проводилось на классах функций GKLS двух видов (Simple and Hard) размерности 2, 3, 4 и 5 из \cite{fio_bib19}.
-Критерием остановки алгоритмов служило попадание точки очередного испытания в $\varepsilon$-окрестность истинного глобального минимума.
-В таблице сравнивается среднее количество итераций, выполненных алгоритмами при решении задач, и ускорение по времени относительно последовательного запуска.
-Распараллеливание  осуществлялось с применением технологии MPI, запуск производился на 8 процессах.
+%Ниже приведены результаты сравнения двух параллельных алгоритмов --- асинхронного GSA и его модификации с использованием decision tree и локального метода Хука-Дживса.
+Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{fio_bib19}.
+Критерием остановки служило попадание точки очередного испытания trial в $\varepsilon$ окрестность истинного глобального минимума.
+%В таблице \ref{tab:2} представлено среднее количество итераций и ускорение по времени относительно последовательного запуска.
+Распараллеливание  осуществлялось с применением  MPI.
+При этом запуск производился на 4 процессах, а следовательно, на каждой итерации вычислялось 3 точки.
+Значение в скобках (если оно есть) показывает количество не решенных задач.
+Это означает, что было достигнут максимум по числу проводимых итераций, но в окрестность глобального минимума точка так и не попала.
+
+Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (GSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented.
+The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard) of dimensionalities 2, 3, 4, and 5 from \cite{fio_bib19}.
+The stop  criterion for the algorithms was the falling of the next trial points into the $\varepsilon$-nearness of  true global minimum.
+In Table \ref{tab:2}, the averaged numbers of iterations executed by the algorithms when solving the  problems and the values of speedup in time relative to the sequential run are compared.
+The  parallelization was performed using MPI technology, the run was performed on 8 processes.
 
 \begin{table}[h!]
-	\caption{Среднее число итераций и среднее число trial, проводимое разными алгоритмами}
+	\caption{Averaged numbers of iterations and averaged numbers of trials executed by different 
+		algorithms}
 	\label{tab:2}
@@
 
-		N & Класс задачи & \multicolumn{2}{c|}{Среднее число итераций} & \multicolumn{2}{c|}{Ускорение} \\ \hline
-		& ~ & АГПА & decision tree & АГПА & decision tree \\ \hline
-		& Simple & 3823,1   & 485,6  & 7,4   & 0,5 \\ \hline
-		2  & Hard & 787,6    & 304,2  & 13,8  & 1,0 \\ \hline
-		& Simple & 5700,5   & 720,1  & 5,6   & 0,1 \\ \hline
-		3  & Hard & 2411,9   & 634,3  & 5,0   & 0,1 \\ \hline
-		& Simple & 31101,5  & 1037,4 & 5,1   & 0,2 \\ \hline
-		4  & Hard & 10418,1  & 816,2  & 6,9   & 0,2 \\ \hline
-		& Simple & 163313,5 & 1204,2 & 4,8   & 1,0 \\ \hline
-		5  & Hard & 27524,5  & 999,4  & 4,6   & 0,2 \\ \hline
+		N & Problem class & \multicolumn{2}{c|}{Averaged number of iterations} & \multicolumn{2}{c|}{Speedup} \\ \hline
+		& ~ & GSAA & Decision tree & GSAA & Decision tree \\ \hline
+		& Simple & 3823,1   & 485,6  & 7,4   & 12,7 \\ \hline
+		2  & Hard & 787,6    & 304,2  & 13,8  & 14,9 \\ \hline
+		& Simple & 5700,5   & 720,1  & 5,6   & 14,2  \\ \hline
+		3  & Hard & 2411,9   & 634,3  & 5,0   & 5,9  \\ \hline
+		& Simple & 31101,5  & 1037,4 & 5,1   & 7,1  \\ \hline
+		4  & Hard & 10418,1  & 816,2  & 6,9   & 5,0  \\ \hline
+		& Simple & 163313,5 & 1204,2 & 4,8   & 4,1 \\ \hline
+		5  & Hard & 27524,5  & 999,4  & 4,6   & 6,3  \\ \hline
 	\end{tabular}
@@
 
-Как видно из таблицы iteration speedup достаточно велико, причем на задачах любой размерности.
+As one can see from Table \ref{tab:2},  the iteration speedup is large enough for the problems of any  dimensionality.
 
-Таким образом, в результате проведенной работы удалось успешно объединить алгоритм глобального поиска с методом локальной оптимизации. 
-В отличие от известных мультистартовых схем решение о запуске локального метода производится с использованием деревьев решений.
-Использование такой комбинации методов позволяет значительно ускорить работу алгоритма.
+So far, as a result of the work done, we succeeded to combine the global search algorithm with a local  optimization method.
+Unlike the known multistart schemes, the decision to run the local method is made using the decision  tree.
+The use of such a combination of methods allows accelerating the work of algorithm considerably.
 
-Параллельный вариант алгоритма сохраняет свойства своего последовательного прототипа, что подтверждается вычислительными экспериментами, во время которых была решена серия из нескольких сотен задач разной размерности.
-Предложенная нами схема позволяет использовать преимущества обоих подходов -- и распараллеливания, и быстрого поиска локальных экстремумов.
+The parallel version of the algorithm preserves the properties of its sequential prototype that was  confirmed by he numerical experiments on solving a series of several hundred problems of various  dimensionalities.
+The scheme we proposed allows utilizing the advantages of both approaches – of the  parallelization as well as of fast search of local extrema.
 
@@
 \bibitem{fio_bib1}
-Ferreiro A.M., Garcia J.A., Lopez-Salas J.G., Vazquez C.
-An efficient implementation of parallel simulated annealing algorithm in GPUs // Journal of global optimization.
-— 2013.
-— Vol.
-57, No.
-3.
-— P.
-863–890.
+M.
+Ferreiro, J.
+A.
+Garcia, J.
+G.
+Lopez-Salas, and C.
+Vazquez, “An efficient implementation of parallel simulated annealing algorithm in gpus,” Journal of global optimization, vol.
+57, no.
+3, pp.
+863–890, 2013.
 
 \bibitem{fio_bib2}
-Garcia-Martinez J.M., Garzon E.M., Ortigosa P.M. A GPU implementation of a hybrid evolutionary algorithm: GPuEGO // Journal of super-computing. — 2014. — Vol. 70, No. 2. — P. 684–695.
+J. M. Garcia-Martinez, E. M. Garzon, and P. M. Ortigosa, “A gpu implementation of a hybrid evolutionary algorithm: Gpuego,” Journal of super-computing, vol. 70, no. 2, pp. 684–695, 2014.
 
 \bibitem{fio_bib3}
-Langdon W.B. Graphics processing units and genetic programming: an overview // Soft Computing. — 2011. — Vol. 15, No 8. — P. 1657–1669.
+W. B. Langdon, “Graphics processing units and genetic programming: an overview,” Soft Computing, vol. 15, no. 8, pp. 1657–1669, 2011.
 
@@
 
-\bibitem{fio_bib8}
-Захарова Е.М., Минашина И.К. Обзор методов многомерной оптимизации. Информационные процессы // том 14 - №3 – 2014 - С. 256-274.
-
-\bibitem{fio_bib9}
-Городецкий С.
-Ю.
-Методические материалы к лабораторной работе «Вычислительные методы поиска локальных минимумов функций», 2001.% URL: http://itmm.unn.ru/files/2016/09/MO_Lab2_LocOpt.pdf (дата обращения: 09.07.2022).
-
-\bibitem{fio_bib10}
-Шефов К.
-С., Степанова М.
-М.
-Реализация и применение параллельного алгоритма глобального поиска минимума к задаче оптимизации параметров молекулярно-динамического потенциала ReaxFF: %URL: http://crm.ics.org.ru/uploads/crmissues/crm_2015_3/15750.pdf, 2015 (дата обращения: 13.10.2020).
-
 \bibitem{fio_bib11}
-Gergel V.P. A global optimization algorithm for multivariate functions with lipschitzian first derivatives / Gergel V.P. // Journal of Global Optimization. – 1997. – Vol. 10, No. 3. – P. 257-281.
+V. P. Gergel, “A global optimization algorithm for multivariate functions with lipschitzian first derivatives,” Journal of Global Optimization, vol. 10, no. 3, pp. 257–281, 1997.
 
@@
 \bibitem{fio_bib13}
-Gaviano, M.
-Software for generation of classes of test functions with known local and global minima for global optimization/ M.
+M.
 Gaviano, D.
@@
 E.
-Kvasov, Y.
+Kvasov, and Y.
 D.
-Sergeyev // ACM Transactions on Mathematical Software.
-– 2003.
-– Vol.
-29.
-– P.
-469-480.
+Sergeyev, “Software for generation of classes of test functions with known local and global minima for global optimization,” ACM Transactions on Mathematical Software, vol.
+29, pp.
+469–480, 2003.
 
 \bibitem{fio_bib14}
-Химмельблау Д. Прикладное нелинейное программирование // М.: Мир - 1975 - 536 с.
+D. Himmelblau, Applied Nonlinear Programming. Mir Publishers, 1975.
 
@@
 \bibitem{fio_bib16}
-OpenCV (Open Source Computer Vision) documentation. %URL: https://docs.opencv.org/4.x/dc/dd6/ml_intro.html (дата обращения: 10.07.2022)
+“Opencv (open source computer vision) documentation,” (accessed: 10 July 2022).
+[Online]. Available: %URL:{https://docs.opencv.org/4.x/dc/dd6/ml_intro.html}
 
 \bibitem{fio_bib17}
-Сергеев, Я.Д. Диагональные методы глобальной оптимизации / Я.Д. Сергеев, Д.Е. Квасов – М.: Физматлит, 2008. – 352 c.
+Ya.D. Sergeev, D.E. Kvasov, Diagonal Global Optimization Methods. Fizmatlit, 2008.
 
 \bibitem{fio_bib18}
-Sysoyev, A.,  Barkalov, K.,  Sovrasov, V.,  Lebedev, I.,  Gergel, V.
-Globalizer – A parallel software system for solving global optimization problems.
-Lecture Notes in Computer Science.
-Volume 10421 LNCS, 2017, Pages 492-499
+Sysoyev, K.
+Barkalov, V.
+Sovrasov, I.
+Lebedev, and V.
+Gergel, “Globalizer – a parallel software system for solving global optimization problems,” Lecture Notes in Computer Science, vol.
+10421, no.
+LNCS, pp.
+492–499, 2017.
 
 \bibitem{fio_bib19}
-Sergeyev, Ya.D.
-Global search based on efficient diagonal partitions and a set of Lipschitz constants / Ya.D.
-Sergeyev, D.E.
-Kvasov // SIAM Journal on Optimization.
-– 2006.
-Vol.
-16, No.
-3.
-– P.
-910–937
+Y. D. Sergeyev and D. E. Kvasov, “Global search based on efficient diagonal partitions and a set of lipschitz constants,” SIAM Journal on Optimization, vol. 16, no. 3, pp. 910–937, 2006.
 
 \bibitem{fio_bib20}
-Стронгин Р.Г. Параллельные вычисления в задачах глобальной оптимизации / Стронгин Р.Г., Гергель В.П., Гришагин В.А., Баркалов К.А. – М.: Издательство Московского университета, 2013. 280 с.
+R.G. Strongin, V.P. Gergel, V.A. Grishagin, K.A. Barkalov , Parallel Computations in Global Optimization Problems. Izd. Mosk. Gos. Univ., 2013.
 

### 2023-01-30 2023/problem-info/artricle.tex
Исправлена схожесть

@@
 \begin{small}
-\textbf{Abstract:} The paper considers the solution of multidimensional global optimization problems using decision trees to identify areas of attraction of local minima.
-The local method is used for local refinement and is designed to speed up the convergence of the algorithm.
-The desired function satisfies the Lipschitz condition with an unknown constant.
-To confirm the theory, computational experiments were presented demonstrating acceleration when solving a series of test problems.
+\textbf{Abstract:} To solve multidimensional global optimization problems, local methods can be used to locally refine and accelerate the convergence of the entire algorithm.
+In order to identify areas of attraction of local minima, decision trees were used in this work.
+For the objective function of the optimization problem, it is only assumed that it satisfies the Lipschitz condition, while the constant is considered unknown.
+To confirm the theory, a number of experiments were given.
+They demonstrate that when using this approach on a series of test problems, it is possible to achieve a significant acceleration in the number of iterations of the algorithm.
 
@@
 \bibliographystyle{IEEEtran}
-\bibliography{articleEng}
+\bibliography{article}
 
@@
 \begin{small}
-\textbf{Аннотация:} В работе рассматривается решение задач многомерной глобальной оптимизации с применением деревьев решений для выявления областей притяжения локальных минимумов.
-Локальный метод используется для локального уточнения и призван ускорить сходимость алгоритма.
+\textbf{Аннотация:} Для решения задач многомерной глобальной оптимизации можно использовать локальные методы для локального уточнения и ускорения сходимости всего адгоритма.
+С целью выявления областей притяжения локальных минимумов, в данной работе применялись деревья решений.
 Для целевой функции задачи оптимизации предполагается только то, что она удовлетворяет условию Липшица, при этом константа считается неизвестной.
-Для подтверждения теории был приведен ряд экспериментов на серии тестовых задач,результаты демонстрируют значительное ускорение по числу итераций алгоритма.
+Для подтверждения теории был приведен ряд экспериментов.
+Они демонстрируют, что при использовании данного подхода на серии тестовых задач можно добиться значительного ускорения по числу итераций алгоритма.
 
@@
 
-При нахождении глобального минимума функции можно использовать разные подходы: от алгоритмов использующих случайный поиск \cite{fio_bib1, fio_bib2, fio_bib3}, до детерминированного, гарантирующего нахождение глобального минимума \cite{.
-fio_bib4, fio_bib5, fio_bib6}.
-В реальных задачах, связанных с глобальной оптимизацией, любое вычисление значения функции является очень трудоемкой задачей, поэтому количество таких операций должно быть уменьшено.
-Это можно сделать путем целенаправленного выбора вариантов в процессе поиска наилучшего решения.
-На этой идее основан алгоритм глобального поиска (AGS).
-В статье мы попробуем объединить АГП и метод локальной оптимизации (метод Хука-Дживса) с помощью деревьев решений, с целью уменьшить количество проводимых итераций, а вместе с этим и количество вычислений целевой функции.
+В прикладных задачах нередко возникает необходимость использования глобальной оптимизации.
+При этом необходимо уменьшить число проводимых в ходе работы алгоритма операций, так как любое вычисление значения функции - трудоемкая задача.
+Это накладывает дополнительные требование на то, какой подход к решению задачи выбрать.
+Самих подходов достаточно много и они реализуют разные идеи: от случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3}, до детерминированного \cite{fio_bib4, fio_bib5, fio_bib6}.
+Однако, учитывая все требования, для наших целей нам лучше всего подходит алгоритм глобального поиска (AGS).
+В процессе поиска он рассматривает не все варианты, а отбирает лучшие по некоторым критериям.
+В статье мы попробуем объединить АГП и метод локальной оптимизации (метод Хука-Дживса), применяя деревья решений.
+В теории, они должны дополнительно уменьшить количество проводимых итераций.
+А это приведет к непосредственному уменьшению вычислений целевой функции в общем.
 
-Алгоритмы локальной оптимизации предназначены для нахождения одного локального экстремума в области определения задачи, в котором целевая функция принимает  минимальное (или максимальное) значение.
-При их построении может использоваться как детерминированный спуск в области экстремальных значений, так и случайный поиск.
-К детерминированным относятся методы нулевого порядка и градиентные методы (первого и второго порядка).
+Методы локальной оптимизации используются для решения более простых задач: когда необходимо найти один из локальных эксремумов функции (точки, в которой целевая функции принимает минимальное или максимальное значение) в области ее определения.
 
@@
 
-Рассмотрим задачу глобальной оптимизации с целевой функцией $\varphi(y)$ на гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1 \le\, i\le\ n \}$ Предположим, что функция удовлетворяет условию Липшица с заранее неизвестной константой $L$.
+Рассмотрим задачу глобальной оптимизации с целевой функцией $\varphi(y)$ на гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1 \le\, i\le\ n \}$.
+Для решения задачи, необходимо выдвинуть дополнительное требование к виду функции - предположим, что она удовлетворяет условию Липшица.
+Значение этой константы нам заранее неизвестно $L$.
 Условие Липшица дает нам  ограниченные вариации значений функций при ограниченных изменениях её аргументов.
@@
 
-При численном решении задачи (\ref{sec:problem}) на основе некоторого числа $k$ вычислений значений целевой функции, происходит построение оценки $y_k^\ast\in D$,  соответствующей понятию близости к точке $y^\ast$ (например, ${ ||y^\ast -y}_k^\ast||\le\ \varepsilon$, где $\varepsilon\geq0$ — заданная точность).
-Для рассматриваемого класса задач предполагается, что оптимизируемая функция $\varphi(y)$ может быть задана алгоритмически как результат работы подпрограммы или библиотеки, а не аналитически в виде формулы.
+При численном решении задачи (\ref{sec:problem}) возможно построить лишь оценку $y_k^\ast\in D$ на основе числа проведенных итераций $k$.
+Чтобы такую оценку можно было использовать как решение задачи, она должна соответствовать выбранному критерию близости к реальной оптимальной точке $y^\ast$.
+В качестве примера можно рассмотреть следующий критерий: ${ ||y^\ast -y}_k^\ast||\le\ \varepsilon$, где $\varepsilon\geq0$ — заданная точность.
+Отдельно отметим оптимизируемую функцию $\varphi(y)$.
+Она не обязана быть представлена в аналитическом виде.
+Это может быть результат работы подпрограммы или библиотеки, что расширяет область ее возможного применения.
 
-По сравнению с другими типами задач оптимизации, задачи многоэкстремальной оптимизации имеют значительно более высокую сложность решения.
-Вычислительные затраты на решение задачи растут экспоненциально с увеличением размерности.
-Глобальная оптимальность является интегральным свойством решаемой задачи и требует исследования во всем пространстве поиска.
-В результате поиск глобального оптимума сводится к построению покрытия (сетки) в области параметров и выборе оптимального значения функции на этой сетке.
+Итак, мы решаем многоэкстремальные многомерные задачи глобальной оптимизации.
+Стоит отметить, что каждый термин в данном описании накладывает свои трудности.
+Задачи многоэкстремальной оптимизации выделяются на фоне других задач оптимизации в силу более высокой сложности решения.
+Глобальная оптимальность требует проведения поиска на всей заданной области, сузить ее заранее невозможно.
+А многомерность ведет к тому, что при росте размерности, трудоемкость поиска решения растет экспоненциально.
+В итоге, необходимо построить покрытие в области параметров (так называемая сетка) и выбирать в этой области оптимальное значение функции.
 
@@
 Алгоритм глобального поиска основан на идее, что минимизируемая функция представляет из себя случайный процесс.
-Решающее правило алгоритма построено так, что следующая итерация выполняется в глобальном минимуме математического ожидания значения функции.
-Найденная точка сохраняется в списке известных значений, после чего итерация повторяется .
-Это продолжается до тех пор, пока не будет достигнут один из выбранных критериев остановки:
+При выполнении каждой новой итерации, найденная точка сохраняется в списке известных значений и  процесс повторяется.
+Главное в подобных алгоритмах - решающее правило и оно в АГП сконструировано таким образом, чтобы каждая последующая итерация проводилась в глобальном минимуме математического ожидания значения функции.
+Более подробно работа алгоритма будет представлена ниже.
+Критерий остановки - это правило, согласно которому алгоритм определяет, необходимо завершить работу или продолжить выполнение.
+В качестве такого правила может быть выбран один из следующих подходов:
 \begin{enumerate}
+	\item Достигнут заданный максимум по числу возможных итераций. 
+ \item Новая точка не будут получена рядом с истинным глобальным минимумом.
 	\item Расстояние между двумя точками не будет меньше заданного значения.
- \item Новая точка не будут получена рядом с истинным глобальным минимумом.
- \item Будет достигнуто заданное максимального числа возможных итераций.
 \end{enumerate}
@@
 сведение многомерной задачи к эквивалентной одномерной) с использованием кривых Пеано.
-Которая позволяют свести исходную многомерную задачу оптимизации в области $D$  к одномерной задаче минимизации на отрезке $[0, 1]$
+Это позволяет не рассматривать задачу на всей области поиска $D$, а ограничиться лишь минимизацией на отрезке $[0, 1]$
 
 \begin{equation*}
-\phi({y(x}^\ast))\ =\ min\{\phi(y(x)):\ x\epsilon[0,\ 1]\}
+	\phi({y(x}^\ast))\ =\ min\{\phi(y(x)):\ x\epsilon[0,\ 1]\}.
 \end{equation*}
-где функция  удовлетворяют уже равномерному условию Гельдера
+Условие Липшица, выдвинутое нами при постановке задачи, при таком переходе в первозданном виде применить нельзя. Однако оно может быть преобразовано в эквивалентное равномерное условие Гельдера
 \begin{equation*}
@@
 
-Поэтому, не ограничивая общности, можно рассматривать минимизацию одномерной функции $f(x)\ =\ \phi(y(x)), \ x\epsilon [0,1]$, удовлетворяющей условию Гельдера.
+Проделанные таким образом преобразования позволяют рассматривать минимизацию одномерной функции $f(x)\ =\ \phi(y(x)), \ x\epsilon [0,1]$. Общность рассуждений при этом нарушена не будет.
 
@@
 В \cite{fio_bib20} приведено обоснование данного способа организации вычислений.
-А в работах  \cite{fio_bib9, fio_bib11} представлены модификации, учитывающие наличие ограничений-неравенств, а также информацию о значение производной целевой функции,
+А в работе  \cite{fio_bib11} представлены модификации, учитывающие наличие ограничений-неравенств, а также информацию о значение производной целевой функции,
 
@@
 
- Метод Хука-Дживса представляет собой комбинацию поиска по образцу и исследующего поиска по направлениям  \cite{fio_bib14, fio_bib15}.
-
-Алгоритм исследующего поиска состоит из следующих шагов: 
+ Метод Хука-Дживса был выбран нами как достаточно эффективный и простой в реализации алгоритм локальной оптимизации.
+Он состоит из двух частей.
+\cite{fio_bib14, fio_bib15}.
+Первая часть - исследующий поиск - включает в себя несколько шагов:
  \begin{enumerate}
@@
 
-Поиск по образцу состоит из выполнения одного шага от полученной базовой точки вдоль линии, соединяющей ее с предыдущей базовой точкой. Величина этого шага определяется задаваемым заранее параметром.
+ Вторая часть - поиск по образцу - состоит из выполнения одного шага от полученной базовой точки вдоль линии, соединяющей ее с предыдущей базовой точкой.
+Величина этого шага определяется задаваемым заранее параметром.
 
@@
 Деревья решений --- это метод автоматического анализа больших массивов данных, который применяется в машинном обучении.
-Деревья решений представляет собой бинарное дерево (дерево, в котором каждый не листовой узел имеет два дочерних узла).
-Его можно использовать как для решения задач классификации, так и для регрессионных задач.
-При решении задач классификации, каждый лист дерева помечается меткой класса, при этом несколько листьев могут иметь одну и ту же метку.
-В случае построение регрессии, каждому листу дерева присваивается константа, как следствие  получаемая функция аппроксимации является кусочно-постоянной.
+Дерево решений представляет собой бинарное дерево (дерево, в котором каждый не листовой узел имеет два дочерних узла).
+Такая структура позволяет использовать подобное дерево в машинном обучении (причем как для решения задач классификации, так и для регрессионных задач) в целях автоатизации анализа большого объема данных.
+Для задач регрессии, каждому листу дерева присваивается константа, как следствие  получаемая функция аппроксимации является кусочно-постоянной.
+А вот для классификации подход иной: каждый лист дерева помечается меткой класса, при этом несколько листьев могут иметь одну и ту же метку.
 
@@
 
-В нашем случае дерево решений строит функцию $\varphi(y)$ кусочно постоянную аппроксимацию$\varphi(y)$  в многомерном пространстве, обозначим как $z' = \psi(y)$ значение вычисленное по decision tree в точке $y$.
+В нашем случае дерево решений строит функцию $\varphi(y)$ кусочно постоянную аппроксимацию $\varphi(y)$  в многомерном пространстве, обозначим как $z' = \psi(y)$ значение вычисленное по decision tree в точке $y$.
 
-При реализации алгоритма  решения задач многомерной глобальной оптимизации с применением деревьев решений для выявления областей притяжения локальных минимумов нами были использованы  деревья решений из библиотеки OpenCV (это библиотека алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения с открытым кодом).
-Подробнее с  деревьями решений можно ознакомиться в \cite{fio_bib16}
+Библиотека алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения с открытым кодом OpenCV предоставляет уже готовые к использованию деревья решений.
+В данной работе, при разработке алгоритма, применяющего деревья решений для выявления областей притяжения локальных минимумов, мы использовали данную библиотеку.
+И, следовательно, все подробности внутренней реализации деревьев решений можно найти в соответствующей документации - \cite{fio_bib16}
 
@@
 В процессе работы АГП необходимо определить, стоит ли использовать текущую точку, в качестве стартовой для локального метода или нет.
-Для этого можно осмотреть соседние точки испытания и если среди них нет ни одной такой, значение функции в которой меньше, чем в текущей, то можно предположить, что мы находимся в области притяжения локального минимума и можно запустить локальный метод, чтобы быстрее сойтись к минимуму.
+Для этого можно осмотреть соседние точки испытания.
+Если среди них есть хоть одна, значение функции в которой больше, чем в текущей, то продолжать работу бессмысленно.
+В противном случае, можно предположить, что мы находимся в области притяжения локального минимума и можно запустить локальный метод, чтобы быстрее сойтись к минимуму.
 Использование проекций многомерных точек на одномерный отрезок после редукции размерности (то, что позволяют сделать кривые Пеано) нам для этих целей не подойдут по нескольким причинам.
 Во-первых, функция после отображения сильно изменяется, один локальный минимум в многомерном пространстве может разделиться на множество минимумов после отображения.
-Во-вторых, после отображения мы можем потерять информацию о взаимном расположении точек исходной функции.
+Во-вторых, после отображения информацию о взаимном расположении точек исходной функции может быть искажена.
 Точки, расположенные близко в многомерном пространстве, могут оказаться на разных концах одномерного отрезка и наоборот).
@@
 При рассмотрении соседей необходимо учитывать следующее: если хоть один из соседей имеет значение функции меньшее, чем в текущей точке, то запуск локального метода не нужен; если какой-то из соседей имеет значение функции такое же, как в текущей, то его также необходимо проверить (найти всех его соседей и проверять их аналогично); и только в том случае, если все значения функций соседей больше (включая соседей равных точек) мы можем запускать локальный метод.
-Модифицированная версия алгоритма глобального поиска выглядит следующим образом:
+Ниже представлен алгоритм глобального поиска, учитывающий описанные ранее модификации:
 
@@
 Все полученные результаты были добавлены в систему globalizer \cite{fio_bib18}.
-Если говорить более конкретно, то была реализована объединенная схема с запуском локальных методов в нужный момент, для чего была внедрена система  построения дерева решений, равномерной сетки и прочих сопутствующих вычислений.
+Если говорить более конкретно, то была внедрена система  построения дерева решений, равномерной сетки и прочих сопутствующих вычислений.
+Использовалась данная система при реализации схемы с запуском локальных методов в нужный момент.
 Ниже приведена блок-схема того, как данная схема была объединена с АГП.
@@
 
-В работе \cite{fio_bib13, fio_bib17} описан GKLS-генератор, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами: количеством локальных минимумов, размерами их областей притяжения, точкой глобального минимума, значением функции в ней и т.п.
-
-\cite{fio_bib13, fio_bib17} описывает генераторы GKLS, которые могут генерировать многоэкстремальные задачи оптимизации с известными свойствами: глобальным минимумом и его значением, количеством локальных минимумов и так далее.
+Работы \cite{fio_bib13, fio_bib17} описывают генераторы GKLS, которые могут генерировать многоэкстремальные задачи оптимизации с известными свойствами: глобальным минимумом и его значением, количеством локальных минимумов и так далее.
 
@@
 
-Ускорение по числу итераций достаточно велико, причем на задачах любой размерности.
-Среднее число точек испытаний тоже, практически везде, уменьшилось, но эти значения могут быть сильно выше, чем среднее число испытаний.
-Это связано с тем, что в точках испытаний учитываются и те точки, которые в процессе своей работы ставил локальный метод, а их может быть довольно много.
+Среднее число точек испытаний, практически везде, уменьшилось, хотя эти значения и могут быть выше, чем среднее число испытаний (ускорение по которым достаточно велико, вне зависимости от сложности задачи и ее размерности).
+Объясняется это тем, что в точках испытаний учитываются и те значения, которые в процессе своей работы ставил локальный метод, а их может быть довольно много.
 
@@
 
-В результате работы нами были успешно совмещены алгоритм глобального поиска с локальным поиском минимума (метод Гука-Джевса).
-Также были проведены вычислительные эксперименты на серии из сотни тестовых задач для экспериментального подтверждения теоретических свойств рассматриваемого алгоритма.
+В результате работы нами были успешно совмещены алгоритм глобального поиска с локальным поиском минимума (метод Хука-Джевса).
+Теоретические свойства рассматриваемого алгоритма удалось подтвердить на практике: серии из сотни тестовых задач показали ускорение по числу проводимых итераций.
 
@@
 
-\paragraph{Лебедев Илья Генадьевич} --- заведующий лабораторией, Нижегородский государственный университет им.
+\paragraph{Лебедев Илья Генадьевич} --- заведующий лабораторией суперкомпьютерных технологий и высокопроизводительных вычислений кафедры математического обеспечения и суперкомпьютерных технологий института информационных технологий, математики и механики, Нижегородский государственный университет им.
 Н.И.
@@
 
-\paragraph{Д.
-И.
-Силенко} --- окончил бакалавриат института информационных технологий, математики и механики Нижегородского государственного университета им.
+\paragraph{Дмитрий Игоревич Силенко} --- окончил бакалавриат института информационных технологий, математики и механики Нижегородского государственного университета им.
 Н.
@@
 Область научных интересов включает методы глобальной и локальной оптимизации, а также технологии параллельных вычислений и параллельные алгоритмы.
-e-mail: \url{dmsi@mail.ru}.
+e-mail: \url{rotor12587@mail.ru}.
 Является автором или соавтором нескольких научных работ в указанных областях.
@@
 
-\paragraph{Ilya~G.~Lebedev} --- head of laboratory, Lobachevsky State University of Nizhny Novgorod.
+\paragraph{Ilya~G.~Lebedev} --- Head of the Laboratory of Supercomputer Technologies and High-Performance Computing, Department of Software and Supercomputer Technologies, Institute of Information Technology, Mathematics and Mechanics, Nizhny Novgorod State University.
+N.I.
+Lobachevsky.
 e-mail: \url{ilya.lebedev@itmm.unn.ru}.
@@
 His research interests are algorithms of global and local optimization, parallel algorithms and parallel technologies.
-e-mail: \url{dmsi@mail.ru}.
+e-mail: \url{rotor12587@mail.ru}.
 He has authored or coauthored several papers in these areas.

### 2023-01-30 2023/PAVT/ParallelDecisionTree.tex
proofread the last section

@@
 % переводить до этого раздела
-\section{Результаты экспериментов}\label{SecR}
+\section{Results and Conclusion}\label{SecR}
 
@@
 Table \ref{tab:1} presents the averaged numbers of iterations executed by respective methods when  solving a series of GKLS problems.
-The sign «>» points to the situation when not all problems are  solved.
-In these cases, the numbers of non-solved problems are given in brackets.
+The sign ``>'' points to the situation when not all problems are  solved.
+In this case, the number of non-solved problems is given in brackets.
 As one can see,  GSA overcomes DIRECT and DIRECTl methods in average number of iterations necessary for solving the problems with the same precision.
@@
 
-Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (GSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented.
-The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard) of dimensionalities 2, 3, 4, and 5 from \cite{fio_bib19}.
+Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (AGSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented.
+The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard from \cite{fio_bib19}) of dimensionalities 2, 3, 4, and 5.
 The stop  criterion for the algorithms was the falling of the next trial points into the $\varepsilon$-nearness of  true global minimum.
-In Table \ref{tab:2}, the averaged numbers of iterations executed by the algorithms when solving the  problems and the values of speedup in time relative to the sequential run are compared.
+In Table \ref{tab:2}, the averaged numbers of iterations executed by the algorithms when solving the  problems and the values of time speedup relative to the sequential run are compared.
 The parallelization was performed using MPI technology, the run was performed on 8 processes.
@@
 		N & Problem class & \multicolumn{2}{c|}{Averaged number of iterations} & \multicolumn{2}{c|}{Speedup} \\ \hline
-		& ~ & GSAA & Decision tree & GSAA & Decision tree \\ \hline
-		& Simple & 3823,1   & 485,6  & 7,4   & 12,7 \\ \hline
-		2  & Hard & 787,6    & 304,2  & 13,8  & 14,9 \\ \hline
-		& Simple & 5700,5   & 720,1  & 5,6   & 14,2  \\ \hline
-		3  & Hard & 2411,9   & 634,3  & 5,0   & 5,9  \\ \hline
-		& Simple & 31101,5  & 1037,4 & 5,1   & 7,1  \\ \hline
-		4  & Hard & 10418,1  & 816,2  & 6,9   & 5,0  \\ \hline
-		& Simple & 163313,5 & 1204,2 & 4,8   & 4,1 \\ \hline
-		5  & Hard & 27524,5  & 999,4  & 4,6   & 6,3  \\ \hline
+		& ~ & AGSA & Decision tree & AGSA & Decision tree \\ \hline
+		& Simple & 3823.1   & 485.6  & 7.4   & 12.7 \\ \hline
+		2  & Hard & 787.6    & 304.2  & 13.8  & 14.9 \\ \hline
+		& Simple & 5700.5   & 720.1  & 5.6   & 14.2  \\ \hline
+		3  & Hard & 2411.9   & 634.3  & 5.0   & 5.9  \\ \hline
+		& Simple & 31101.5  & 1037.4 & 5.1   & 7.1  \\ \hline
+		4  & Hard & 10418.1  & 816.2  & 6.9   & 5.0  \\ \hline
+		& Simple & 163313.5 & 1204.2 & 4.8   & 4.1 \\ \hline
+		5  & Hard & 27524.5  & 999.4  & 4.6   & 6.3  \\ \hline
 	\end{tabular}
@@
 Unlike the known multistart schemes, the decision to run the local method is made using the decision  tree.
-The use of such a combination of methods allows accelerating the work of algorithm considerably.
+The use of such a combination of methods allows considerably accelerating the algorithm.
 
 The parallel version of the algorithm preserves the properties of its sequential prototype that was  confirmed by he numerical experiments on solving a series of several hundred problems of various  dimensionalities.
-The scheme we proposed allows utilizing the advantages of both approaches – of the  parallelization as well as of fast search of local extrema.
+The proposed scheme allows us utilizing the advantages of both parallelization as well as fast search of local extrema.
 

### 2023-01-30 2023/PAVT/ParallelDecisionTree.tex
Поправил список литературы.

--- 2023-01-30 2023/PAVT/ParallelDecisionTree.tex
@@
 \usepackage{hyperref}
-\usepackage[english, russian]{babel} 
+\usepackage[english]{babel} 
 \usepackage{amsmath} 
@@
 Such problems often arise in the cases, when it is necessary to select the values  of parameters of the mathematical model being investigated, in which the simulation results best fit the experimental data.
-For solving the problems of this class, many algorithms are known: from the metaheuristic algorithms, based on the  idea of random search \cite{fio_bib1,fio_bib2,fio_bib3}, to the deterministic algorithms guarantying convergence to the global minimum \cite{fio_bib4,fio_bib5,fio_bib6}.
+For solving the problems of this class, many algorithms are known: from the metaheuristic algorithms, based on the  idea of random search \cite{Ferreiro2013,Garcia2014,Langdon2011}, to the deterministic algorithms guarantying convergence to the global minimum \cite{Evtushenko2009,He2008,Paulavicius2011}.
 
@@
 It  can be achieved by intentional choice of variants in the course of search for the optimal solution cutting  off unpromising search subdomains and investigating only these ones, in which the solution of the  problem can be found.
-The global search algorithm (GSA) is based on this idea \cite{fio_bib7}.
+The global search algorithm (GSA) is based on this idea.
 In  the present work, we tried to combine GSA and a local optimization method (Hooke-Jeeves pattern search method) to reduce the number of trials executed.
@@
 The local optimization algorithms are intended for determining only one of local extrema within a set  of feasible solutions, in which the objective function takes the extremal (maximum or minimum) value.
-Among the local methods,  the zero-order methods and the gradient methods (of the 1\textsuperscript{st} order and of the  2\textsuperscript{nd} one) are traditionally distinguished \cite{fio_bib8,fio_bib9}.
+Among the local methods,  the zero-order methods and the gradient methods (of the 1\textsuperscript{st} order and of the  2\textsuperscript{nd} one) are traditionally distinguished.
 Note that the application of the first-order methods (not to mention the second one) in the problems  with the ``black-box'' type functions is difficult since the problem of numerical estimating the gradient arises here.
@@
 
-The substantiation of this method of organization of computations see in \cite{fio_bib20}.
-The  modifications taking into account the presence of inequality constraints as well as the information on the derivative of the objective function are presented in \cite{fio_bib12, fio_bib9,  fio_bib11}.
+The substantiation of this method of organization of computations see in \cite{Strongin2013}.
+The  modifications taking into account the presence of inequality constraints as well as the information on the derivative of the objective function are presented in \cite{Barkalov2002,Gergel1997}.
 
@@
 The Hooke--Jeeves pattern search algorithm belongs to the class of zero-order methods.
-Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{fio_bib14, fio_bib15}.
+Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{NelderMead,Himmelblau75}.
 
@@
 OpenCV is an open source library of the algorithms of computer vision, image processing, and  general-purpose numerical algorithms.
-More details on the decision tree can be found  in \cite{fio_bib16}.
+More details on the decision tree can be found  in \cite{Opencv}.
 
@@
 % переводить до этого раздела
-\section{Результаты экспериментов}\label{SecR}
+\section{Numerical Experiments}\label{SecR}
 
@@
 %Ниже приведены результаты сравнения двух параллельных алгоритмов --- асинхронного GSA и его модификации с использованием decision tree и локального метода Хука-Дживса.
-Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{fio_bib19}.
+Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{Sergeyev2006}.
 Критерием остановки служило попадание точки очередного испытания trial в $\varepsilon$ окрестность истинного глобального минимума.
@@
 Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (GSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented.
-The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard) of dimensionalities 2, 3, 4, and 5 from \cite{fio_bib19}.
+The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard) of dimensionalities 2, 3, 4, and 5 from \cite{Sergeyev2006}.
 The stop  criterion for the algorithms was the falling of the next trial points into the $\varepsilon$-nearness of  true global minimum.
@@
 \bibliographystyle{spmpsci}
-\begin{thebibliography}{6}
-%
+\bibliography{bibliography}{}
 
-\bibitem{fio_bib1}
-M.
-Ferreiro, J.
-A.
-Garcia, J.
-G.
-Lopez-Salas, and C.
-Vazquez, “An efficient implementation of parallel simulated annealing algorithm in gpus,” Journal of global optimization, vol.
-57, no.
-3, pp.
-863–890, 2013.
-
-\bibitem{fio_bib2}
-J. M. Garcia-Martinez, E. M. Garzon, and P. M. Ortigosa, “A gpu implementation of a hybrid evolutionary algorithm: Gpuego,” Journal of super-computing, vol. 70, no. 2, pp. 684–695, 2014.
-
-\bibitem{fio_bib3}
-W. B. Langdon, “Graphics processing units and genetic programming: an overview,” Soft Computing, vol. 15, no. 8, pp. 1657–1669, 2011.
-
-\bibitem{fio_bib4}
-Евтушенко Ю.Г., Малкова В.У., Станевичюс А.А. Параллельный поиск глобального экстремума функций многих переменных // Ж. вычисл. матем. и матем. физ. — 2009. — Т. 49, №2. — С. 255–269.
-
-\bibitem{fio_bib5}
-He J., Verstak A., Watson L.T., Sosonkina M. Design and implementation of a mas-sively parallel version of DIRECT // Computational optimization and applications. — 2008. — Vol. 40, No. 2. — P. 217–245
-
-\bibitem{fio_bib6}
-Paulavicius R., Žilinskas J., Grothey A.
-Parallel branch and bound for global optimiza-tion with combination of Lipschitz bounds // Optimization methods and software.
-— 2011.
-— Vol.
-26, No.
-3.
-— P.
-487–498.
-
-\bibitem{fio_bib7}
-Глобальная оптимизация: приложения и вычислительная сложность. %URL: http://hpc-education.unn.ru/ru/globopt/глобальная-оптимизация-приложения-и (дата обращения: 10.10.2020).
-
-\bibitem{fio_bib11}
-V. P. Gergel, “A global optimization algorithm for multivariate functions with lipschitzian first derivatives,” Journal of Global Optimization, vol. 10, no. 3, pp. 257–281, 1997.
-
-\bibitem{fio_bib12}
-Barkalov K.A.
-A global optimization technique with an adaptive order of checking for constraints / Barkalov K.A., Strongin R.G.
-// Computational Mathematics and Mathematical Physics.
-– 2002.
-– Vol.
-42, No.
-9.
-– P.
-1289-1300.
-
-\bibitem{fio_bib13}
-M.
-Gaviano, D.
-Lera, D.
-E.
-Kvasov, and Y.
-D.
-Sergeyev, “Software for generation of classes of test functions with known local and global minima for global optimization,” ACM Transactions on Mathematical Software, vol.
-29, pp.
-469–480, 2003.
-
-\bibitem{fio_bib14}
-D. Himmelblau, Applied Nonlinear Programming. Mir Publishers, 1975.
-
-\bibitem{fio_bib15}
-Nelder J., Mead R. A simplex method for function minimization // Computer Journal - 7(4) – 1965 - P. 308-313.
-
-\bibitem{fio_bib16}
-“Opencv (open source computer vision) documentation,” (accessed: 10 July 2022).
-[Online]. Available: %URL:{https://docs.opencv.org/4.x/dc/dd6/ml_intro.html}
-
-\bibitem{fio_bib17}
-Ya.D. Sergeev, D.E. Kvasov, Diagonal Global Optimization Methods. Fizmatlit, 2008.
-
-\bibitem{fio_bib18}
-Sysoyev, K.
-Barkalov, V.
-Sovrasov, I.
-Lebedev, and V.
-Gergel, “Globalizer – a parallel software system for solving global optimization problems,” Lecture Notes in Computer Science, vol.
-10421, no.
-LNCS, pp.
-492–499, 2017.
-
-\bibitem{fio_bib19}
-Y. D. Sergeyev and D. E. Kvasov, “Global search based on efficient diagonal partitions and a set of lipschitz constants,” SIAM Journal on Optimization, vol. 16, no. 3, pp. 910–937, 2006.
-
-\bibitem{fio_bib20}
-R.G. Strongin, V.P. Gergel, V.A. Grishagin, K.A. Barkalov , Parallel Computations in Global Optimization Problems. Izd. Mosk. Gos. Univ., 2013.
-
-\end{thebibliography}
 \end{document}

### 2023-01-30 2023/PAVT/ParallelDecisionTree.tex

@@
 The Hooke--Jeeves pattern search algorithm belongs to the class of zero-order methods.
-Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{NelderMead,Himmelblau75}.
+Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{Himmelblau75}.
 
@@
 
-Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (GSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented.
-The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard) of dimensionalities 2, 3, 4, and 5 from \cite{Sergeyev2006}.
+Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (AGSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented.
+The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard from \cite{fio_bib19}) of dimensionalities 2, 3, 4, and 5.
 The stop  criterion for the algorithms was the falling of the next trial points into the $\varepsilon$-nearness of  true global minimum.
-In Table \ref{tab:2}, the averaged numbers of iterations executed by the algorithms when solving the  problems and the values of speedup in time relative to the sequential run are compared.
+In Table \ref{tab:2}, the averaged numbers of iterations executed by the algorithms when solving the  problems and the values of time speedup relative to the sequential run are compared.
 The parallelization was performed using MPI technology, the run was performed on 8 processes.

### 2023-01-30 2023/PAVT/ParallelDecisionTree.tex
reduce similarity on 2%

@@
 \end{displaymath}
-This suggestion can be interpreted (with respect to the applied problems ) as the reflection of limited  power generating the changes in the system being simulated.
+This suggestion can be considered (with respect to the real-world problems) as the reflection of limited  power generating the changes in the system being simulated.
 
@@
 
-The solving of the multiextremal optimization problems is much more computation-costly  as compared  to other types of optimization problems since the global optimum is an integral characteristic of the  problem being solved and requires investigating the whole search domain.
-As a result, the search of the  global optimum is reduced to constructing some coverage (grid) in the search domain and the choice of  the best function value on this grid.
+The solving of the global optimization problems is much more computationally expensive as compared to the local optimization problems since finding the global optimum requires exploration of the whole search domain.
+As a result, the search of the  global optimum is reduced to constructing some coverage in the search domain and choosing the best function value on this grid.
 When using the uniform grids, the computation costs of solving the  problem grow exponentially with increasing dimensionality.
@@
 
-Let us assume that the master process computes one point of the next trial at every iteration and sends  it to the worker process to execute the trial.
-Note that executing a trial by the worker process in  the applied optimization problems is much more computation costly than the choice of a new trial point  by the master process.
+Let us assume that the master process computes one point of the next trial at every iteration and sends  it to the slave process to execute the trial.
+Note that executing a trial by the slave process in  the applied optimization problems is much more computation costly than the choice of a new trial point  by the master process.
 Therefore, the idle time of the processes will be negligible.
-In this  case (unlike the synchronous parallel algorithms) total number of trials executed by each worker  process will depend on the computation costs for executing particular trials and cannot be estimated in  advance.
-When describing the parallel algorithm, let us assume that we have $p+1$ computational  processes at our disposal: one master and $p$ workers.
+In this  case (unlike the synchronous parallel algorithms) total number of trials executed by each slave  process will depend on the computation costs for executing particular trials and cannot be estimated in  advance.
+When describing the parallel algorithm, let us assume that we have one master and $p$ slave processes at our disposal.
 
-In the beginning of search, the master process (assume it to have the identifier $0$) initiates parallel  execution of $p$ trials in $p$ different points of the search domain.
-Two of these points are the boundary ones, the rest are the internal points, i.e.
-at the points  $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$ where $x^1=0,x^p=1,x^i\in\left(0,1\right),i=2,\ldots,p-1$.
+At the initial phase the master process (assume it to have the identifier $0$) starts $p$ trials in parallel at $p$ different points $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$.
+The inverse images of two points are the boundary ones $x^1=0,x^p=1$; the rest are the internal points $x^i\in\left(0,1\right),i=2,\ldots,p-1$.
 
-Now assume $k\geq0$ trials to be completed, and the worker processes are performing the trials at the points $y\left(x^{k+1}\right),\;y\left(x^{k+2}\right),\ldots,\;y\left(x^{k+p}\right)$.
+Now assume $k\geq0$ trials to be completed, and the slave processes are performing the trials at the points $y\left(x^{k+1}\right),\;y\left(x^{k+2}\right),\ldots,\;y\left(x^{k+p}\right)$.
 
-Each worker process having completed the computing of the function at some time moment sends  the computed value to the master process.
-In turn, the master process selects a new point  $x^{k+p+1}$ for the worker process according to the rules described below.
-Note that in this case  we will have a set of preimages of the trial points  $I_k=\left\{x^{k+1},x^{k+2},\ldots,x^{k+p}\right\}$, in which the trials have been already started but haven't been completed yet.
+Each slave process having completed the computing of the function at some time moment sends  the computed value to the master process.
+Next, the master process selects a new point  $x^{k+p+1}$ for the slave process according to the rules provided below.
+Note that in this case it will be a set of inverse images of the trial points  $I_k=\left\{x^{k+1},x^{k+2},\ldots,x^{k+p}\right\}$, in which the trials have been already initiated but haven't been completed yet.
 
@@
 
-\begin{enumerate} \item Arrange in the increasing order (by the lower indices) the set of preimages of the trial points
+\begin{enumerate} \item Arrange in the increasing order (by the lower indices) the set of inverse images of the trial points
 	\begin{displaymath} 
@@
 	where $ z_i=\varphi\left(y\left(x_i\right)\right)$ if $x_i\notin I_k,\;1\le i\le k+p$.
-The values $z_i$ at the points $x_i\in I_k$ are undefined since the trials at the points $x_i\in I_k$ haven't been completed yet.
-If the value of $M$ equals to $0$, set $M=1$.
+The values $z_i$ at the points $x_i\in I_k$ are not defined as far as the trials at these points haven't been completed yet.
+If $M=0$, then set $M=1$.
 
@@
 
-	\item Select the interval $\left[x_{t-1},x_t\right]$, which corresponds to the highest characteristic, i.e.
+	\item Find the interval $\left[x_{t-1},x_t\right]$, which corresponds to the highest characteristic, i.e.
 	\begin{displaymath} 
@@
 
-	\item Define a new trial point $y^{k+p+1}=y\left(x^{k+p+1}\right)$, the preimage of which is  $x^{k+p+1}\in\left(x_{t-1},x_t\right)$, according to the formula 
+	\item Compute a new trial point $y^{k+p+1}=y\left(x^{k+p+1}\right)$, the inverse image of which is  $x^{k+p+1}\in\left(x_{t-1},x_t\right)$, according to the formula 
 \begin{displaymath}
@@
 
-Upon computing the next trial point, the master process adds it to the set $I_k$ and sends to the  worker process, which initiates the new trial in this point.
-The master process stops the algorithm, if one of the two conditions is satisfied: $\Delta_t<$ or  $k+p>K_{max}$.
-The real value $\epsilon>0$ and the integer value $K_{max}>0$ are the parameters of the algorithm  and correspond to the accuracy of the search and the maximum number of trials, respectively.
+After the next trial point is computed, the master process includes it to the set $I_k$ and sends to the  slave process, which starts the new trial at this point.
+The master process stops the algorithm, if at least one of the two conditions is satisfied: $\Delta_t<$ or  $k+p>K_{max}$.
+The real value $\epsilon>0$ and the integer value $K_{max}>0$ are the algorithm parameters.
+They correspond to the accuracy of the search and the maximum number of trials, respectively.
 

### 2023-01-30 2023/problem-info/artricle.tex
Поправлен английский текст.

@@
 
-\title{Global optimization algorithm using decision trees to find local extrema \thanks{The work was supported by the oneAPI Center of Excellence, by the Ministry of Science and Higher Education of the Russian Federation (project no.
+\title{Global optimization algorithm that uses decision trees to find local extrema \thanks{The work was supported by the Ministry of Science and Higher Education of the Russian Federation (project no.
 0729-2020-0055), and by the Research and Education Mathematical Center (project no.
-075-02-2021-1394).}}
+075-02-2022-883).}}
 
@@
 \begin{small}
-\textbf{Abstract:} To solve multidimensional global optimization problems, local methods can be used to locally refine and accelerate the convergence of the entire algorithm.
-In order to identify areas of attraction of local minima, decision trees were used in this work.
-For the objective function of the optimization problem, it is only assumed that it satisfies the Lipschitz condition, while the constant is considered unknown.
-To confirm the theory, a number of experiments were given.
-They demonstrate that when using this approach on a series of test problems, it is possible to achieve a significant acceleration in the number of iterations of the algorithm.
+\textbf{Abstract:} The paper considers the algorithms for solving the multidimensional global optimization problems using decision tree to reveal the attraction regions of the local minima.
+We suppose, that the target function is defined as a “black box” and satisfied Lipschitz condition with unknown constant.
+We propose a method for selecting the local extrema neighborhood of the target function based on analysis of accumulated search information using machine learning methods.
+This allows us to make a decision to run a local method, which can speed up the convergence of the algorithm.
+The proposition was confirmed by the results of numerical experiments demonstrating the speedup when solving a series of test problems.
 
@@
 \emptythanks
-\title{Алгоритм глобальной оптимизации, использующий деревья решений для выявления локальных экстремумов\thanks{Работа выполнена при финансовой поддержке Министерства науки и высшего образования РФ (проект № 0729-2020-0055) и научно-образовательного математического центра “Математика технологий будущего” (соглашение № 075-02-2021-1394).}}
+\title{Алгоритм глобальной оптимизации, использующий деревья решений для выявления локальных экстремумов\thanks{Работа выполнена при финансовой поддержке Министерства науки и высшего образования РФ (проект № 0729-2020-0055) и научно-образовательного математического центра “Математика технологий будущего” (соглашение № 075-02-2022-883).}}
 
@@
 \begin{small}
-\textbf{Аннотация:} Для решения задач многомерной глобальной оптимизации можно использовать локальные методы для локального уточнения и ускорения сходимости всего адгоритма.
-С целью выявления областей притяжения локальных минимумов, в данной работе применялись деревья решений.
-Для целевой функции задачи оптимизации предполагается только то, что она удовлетворяет условию Липшица, при этом константа считается неизвестной.
-Для подтверждения теории был приведен ряд экспериментов.
-Они демонстрируют, что при использовании данного подхода на серии тестовых задач можно добиться значительного ускорения по числу итераций алгоритма.
+\textbf{Аннотация:} В работе рассматривается решение задач многомерной глобальной оптимизации с применением деревьев решений для выявления областей притяжения локальных минимумов.
+Предполагается, что целевая функция задачи задана как «черный ящик» и удовлетворяет условию Липшица с априори неизвестной константой.
+Мы предлагаем способ выделения окрестностей локальных экстремумов целевой функции на основе анализа накопленной поисковой информации средствами машинного обучения.
+Это позволяет принять решение о запуске локального метода для уточнения значения функции в локальном минимуме, что может ускорить сходимость алгоритма.
+Выдвинутое предположение подтверждается результатами вычислительных экспериментов, демонстрирующих ускорение при решении серии тестовых задач.
 
@@
 Область научных интересов включает методы глобальной и локальной оптимизации, а также технологии параллельных вычислений и параллельные алгоритмы.
-e-mail: \url{rotor12587@mail.ru}.
-Является автором или соавтором нескольких научных работ в указанных областях.
+e-mail: \url{silenko@itmm.unn.ru}.
+Является соавтором нескольких научных работ в указанных областях.
 
@@
 
-\paragraph{Ilya~G.~Lebedev} --- Head of the Laboratory of Supercomputer Technologies and High-Performance Computing, Department of Software and Supercomputer Technologies, Institute of Information Technology, Mathematics and Mechanics, Nizhny Novgorod State University.
+\paragraph{Ilya~G.~Lebedev} --- Head of the Laboratory of Supercomputer Technologies and High-Performance Computing, Department of Mathematical Software and Supercomputing Technologies, Institute of Information Technology, Mathematics and Mechanics, Nizhny Novgorod State University.
 N.I.
@@
 His research interests include algorithms of global and local optimization, parallel algorithms, GPU programming, CUDA.
-He has authored or coauthored more than 30 papers in these areas.
+He is the author or coauthor more than 30 papers in these areas.
 
@@
 He is currently a master student of the Institute of Information Technology, Mathematics and Mechanics.
-His research interests are algorithms of global and local optimization, parallel algorithms and parallel technologies.
-e-mail: \url{rotor12587@mail.ru}.
-He has authored or coauthored several papers in these areas.
+His research interests are algorithms of global and local optimization, parallel algorithms and parallel computing.
+e-mail: \url{silenko@itmm.unn.ru}.
+He is coauthor of the several papers in these areas.
 

### 2023-01-31 2023/PAVT/ParallelDecisionTree.tex
Заменены картинки
Добалены картинки.

@@
 		\begin{minipage}[h]{0.8\linewidth} 
-			\includegraphics[width=1\linewidth]{figure/fig1.png} 
+			\includegraphics[width=1\linewidth]{figure/fig1.pdf} 
 			\caption{An example of iterations of the Hooke-Jeeves algorithm} %%  подпись к рисунку 
@@
 
-%Fig. \ref{fig:fig2} presents a contour plot of an objective function and the trial points; Fig.  \ref{fig:fig2_2} presents corresponding decision tree.
+Fig. \ref{fig:fig2} presents a contour plot of an objective function and the trial points; Fig.  \ref{fig:fig2_2} presents corresponding decision tree.
+
+\begin{figure}[ht!]
+	\begin{center}
+		\begin{minipage}[h]{0.9\linewidth}
+			\includegraphics[width=1\linewidth]{figure/fig4.png}
+			\caption{A function from the GKLS generator before using Peano curves} %% подпись к рисунку
+			\label{fig:fig2}
+		\end{minipage}
+	\end{center}
+\end{figure}
+
+\begin{figure}[ht!]
+	\begin{center}
+		\begin{minipage}[h]{0.9\linewidth}
+			\includegraphics[width=1\linewidth]{figure/fig5.png}
+			\caption{A figure with a two-dimensional decision tree} %% подпись к рисунку
+			\label{fig:fig2_2}
+		\end{minipage}
+	\end{center}
+\end{figure}
 
@@
 	\begin{center} 
-				\begin{minipage}[h]{0.9\linewidth} 			\includegraphics[width=1\linewidth]{figure/fig3.png}
+				\begin{minipage}[h]{0.9\linewidth} 			\includegraphics[width=1\linewidth]{figure/fig3.pdf}
 					 			\caption{Block scheme of combined GSA and Hooke-Jeeves method using the  decision tree} %% подпись к рисунку 			\label{fig:fig3}

### 2023-01-31 2023/PAVT/ParallelDecisionTree.tex
Поправлен список литературы.

@@
 The decision rules of the grid building algorithm are constructed in such a way that the next  grid point corresponds to the global minimum point of the mathematical expectation of the function  values.
-This point is stored in the list of known values, and the iterations are repeated until the stop criteria is satisfied: either the distance between the trial points becomes less than a  predefined value or a preset maximum number of iterations is achieved \cite{fio_bib10}.
+This point is stored in the list of known values, and the iterations are repeated until the stop criteria is satisfied: either the distance between the trial points becomes less than a  predefined value or a preset maximum number of iterations is achieved.
 
@@
 
-The substantiation of this method of organization of computations see in \cite{Strongin2013}.
+The substantiation of this method of organization of computations see in \cite{Strongin2000,Barkalov2016}.
 The  modifications taking into account the presence of inequality constraints as well as the information on the derivative of the objective function are presented in \cite{Barkalov2002,Gergel1997}.
@@
 The Hooke--Jeeves pattern search algorithm belongs to the class of zero-order methods.
-Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{Himmelblau75}.
+Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{Himmelblau72}.
 
@@
 OpenCV is an open source library of the algorithms of computer vision, image processing, and  general-purpose numerical algorithms.
-More details on the decision tree can be found  in \cite{Opencv}.
+More details on the decision tree can be found  in \cite{Brahmbhatt2013}.
 
@@
 Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (AGSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented.
-The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard from \cite{fio_bib19}) of dimensionalities 2, 3, 4, and 5.
+The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard from \cite{Sergeyev2006}) of dimensionalities 2, 3, 4, and 5.
 The stop  criterion for the algorithms was the falling of the next trial points into the $\varepsilon$-nearness of  true global minimum.

### 2023-01-31 2023/PAVT/ParallelDecisionTree.tex
add some references

@@
 It  can be achieved by intentional choice of variants in the course of search for the optimal solution cutting  off unpromising search subdomains and investigating only these ones, in which the solution of the  problem can be found.
-The global search algorithm (GSA) is based on this idea.
-In  the present work, we tried to combine GSA and a local optimization method (Hooke-Jeeves pattern search method) to reduce the number of trials executed.
+The global search algorithm (GSA) is based on this idea \cite{Strongin2000}.
+In  the present work, we tried to combine GSA and a local optimization method (Hooke-Jeeves pattern search method \cite{HookJeeves}) to reduce the number of trials executed.
 The decision on the run of the local method will be made using a decision tree.
@@
 The local optimization algorithms are intended for determining only one of local extrema within a set  of feasible solutions, in which the objective function takes the extremal (maximum or minimum) value.
-Among the local methods,  the zero-order methods and the gradient methods (of the 1\textsuperscript{st} order and of the  2\textsuperscript{nd} one) are traditionally distinguished.
-Note that the application of the first-order methods (not to mention the second one) in the problems  with the ``black-box'' type functions is difficult since the problem of numerical estimating the gradient arises here.
+Among the local methods,  the zero-order methods and the gradient methods (of the 1\textsuperscript{st} order and of the  2\textsuperscript{nd} one) are traditionally distinguished \cite{Nocedal}.
+Note that the application of the first-order methods (not to mention the second one) in the problems  with the ``black-box'' type functions is difficult since the problem of numerical estimating the gradient arises here \cite{Kelley}.
 Therefore, further we will consider only the zero-order methods.

### 2023-01-31 2023/problem-info/artricle.tex
Обновлен текст, добавлены рисунки.

@@
 \title{Global optimization algorithm that uses decision trees to find local extrema \thanks{The work was supported by the Ministry of Science and Higher Education of the Russian Federation (project no.
-0729-2020-0055), and by the Research and Education Mathematical Center (project no.
+FSWR-2023-0034), and by the Research and Education Mathematical Center (project no.
 075-02-2022-883).}}
@@
 % References
-\renewcommand{\refname}{References}
-\bibliographystyle{IEEEtran}
-\bibliography{article}
+%\renewcommand{\refname}{References}
+%\bibliographystyle{IEEEtran}
+%\bibliography{articleEng}
 
@@
 \emptythanks
-\title{Алгоритм глобальной оптимизации, использующий деревья решений для выявления локальных экстремумов\thanks{Работа выполнена при финансовой поддержке Министерства науки и высшего образования РФ (проект № 0729-2020-0055) и научно-образовательного математического центра “Математика технологий будущего” (соглашение № 075-02-2022-883).}}
+\title{Алгоритм глобальной оптимизации, использующий деревья решений для выявления локальных экстремумов\thanks{Работа выполнена при финансовой поддержке Министерства науки и высшего образования РФ (проект № FSWR-2023-0034) и научно-образовательного математического центра “Математика технологий будущего” (соглашение № 075-02-2022-883).}}
 
@@
 Это накладывает дополнительные требование на то, какой подход к решению задачи выбрать.
-Самих подходов достаточно много и они реализуют разные идеи: от случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3}, до детерминированного \cite{fio_bib4, fio_bib5, fio_bib6}.
+Самих подходов достаточно много, и они реализуют разные идеи: от случайного поиска \cite{fio_bib1, fio_bib2, fio_bib3}, до детерминированного \cite{fio_bib4, fio_bib5, fio_bib6}.
 Однако, учитывая все требования, для наших целей нам лучше всего подходит алгоритм глобального поиска (AGS).
@@
 
-Методы локальной оптимизации используются для решения более простых задач: когда необходимо найти один из локальных эксремумов функции (точки, в которой целевая функции принимает минимальное или максимальное значение) в области ее определения.
+Методы локальной оптимизации используются для решения более простых задач: когда необходимо найти один из локальных экстремумов функции (точки, в которой целевая функции принимает минимальное или максимальное значение) в области ее определения.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%
 \begin{equation} \label{sec:problem}   
@@
 При выполнении каждой новой итерации, найденная точка сохраняется в списке известных значений и  процесс повторяется.
-Главное в подобных алгоритмах - решающее правило и оно в АГП сконструировано таким образом, чтобы каждая последующая итерация проводилась в глобальном минимуме математического ожидания значения функции.
+Главное в подобных алгоритмах --- решающее правило и оно в АГП сконструировано таким образом, чтобы каждая последующая итерация проводилась в глобальном минимуме математического ожидания значения функции.
 Более подробно работа алгоритма будет представлена ниже.
-Критерий остановки - это правило, согласно которому алгоритм определяет, необходимо завершить работу или продолжить выполнение.
+Критерий остановки --- это правило, согласно которому алгоритм определяет, необходимо завершить работу или продолжить выполнение.
 В качестве такого правила может быть выбран один из следующих подходов:
@@
 \begin{equation*}
-	\phi({y(x}^\ast))\ =\ min\{\phi(y(x)):\ x\epsilon[0,\ 1]\}.
+	\phi({y(x}^\ast))\ =\ min\{\phi(y(x)):\ x\in[0,\ 1]\}.
 \end{equation*}
@@
 \begin{equation*}
-	\left|\phi (y \left(x_1\right))- \phi (y \left(x_2\right)\right )|\le\ H\left|x_1-x_2\right|^\frac{1}{N},\ x_1,\ x_2\epsilon[0,1]		
+	\left|\phi (y \left(x_1\right))- \phi (y \left(x_2\right)\right )|\le\ H\left|x_1-x_2\right|^\frac{1}{N},\ x_1,\ x_2\in[0,1]		
 \end{equation*}
 
-Проделанные таким образом преобразования позволяют рассматривать минимизацию одномерной функции $f(x)\ =\ \phi(y(x)), \ x\epsilon [0,1]$. Общность рассуждений при этом нарушена не будет.
+Проделанные таким образом преобразования позволяют рассматривать минимизацию одномерной функции $f(x)\ =\ \phi(y(x)), \ x\in [0,1]$. Общность рассуждений при этом нарушена не будет.
 
@@
 	\item  Перенумеровать (нижним индексом) точки ранее проведенных испытаний $x^i, 1\leq i\leq k$, а также граничные точки отрезка [0,1] в порядке возрастания координаты:
- \begin{equation}
+ \begin{displaymath}
 		\label{agp1_sort}
 	0=x_0<\ x_1<\ ...\ <x_{k+1}=1.
-	\end{equation}
+	\end{displaymath}
 	и сопоставить им значения $z_i=f(x_i)$.
@@
 	\item  Вычислить текущие нижние оценки $M$ неизвестной константы Гельдера $H$:
- \begin{equation}
+ \begin{displaymath}
 		\label{agp2_mu}
 	\mu=max\left\{\frac{|z_i-z_{i-1}|}{{{(x}_i-x_{i-1})}^{1/N}},\ i=1,\ldots,k\right\},\ M=\ \left\{\begin{matrix}r\mu,\ \mu>0,\\1,\ \mu=0,\\\end{matrix}\right.\
-	\end{equation}
+	\end{displaymath}
 	где $r>1$ --- параметр алгоритма.
@@
 	\item  Для всех интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить значение $R(i)$, называемое \textit{характеристикой} интервала, в соответствии с формулами
-	\begin{equation}
+	\begin{displaymath}
 		\label{agp3_R1}
 		R(1)=2\Delta_1-4\dfrac{z_1}{M}, \; R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
-	\end{equation}
-	\begin{equation}
+	\end{displaymath}
+	\begin{displaymath}
 		\label{agp3_Ri}
 		R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1,
-	\end{equation}
+	\end{displaymath}
 	где \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
@@
 	\item   Упорядочить характеристики $R\left(i\right),\ 1\leq i \leq k+1,$ в порядке невозрастания
+
+	%%%%%%%%%%%%%%%%%%%%%
 	\begin{equation}
@@
 	\item  В выбранном интервале вычислить точку $x^{k+1}$, в соответствии с формулами
-	\begin{equation}
+	\begin{displaymath}
 		\label{agp5_x1}
 	x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\ t_j=1,\ t_j=k+1,
-	\end{equation}	
-	\begin{equation}
+	\end{displaymath}	
+	\begin{displaymath}
 		\label{agp4_xi}	
 	x^{k+1}=\frac{x_{t_j}+x_{t_j-1}}{2}-sign\left(z_{t_j}-z_{t_j-1}\right)\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N,\ 1<t_j<k+1.
-	\end{equation}
+	\end{displaymath}
 
@@
 Результатом работы АГП при решении рассматриваемой задачи, является оценка глобального оптимума:
-\begin{equation}
+\begin{displaymath}
 	f_k^*=\min_{1\leq i \leq k}f(x_i), \; x_k^*=arg \min_{1\leq i \leq k}f(x_i).
-\end{equation}
+\end{displaymath}
 
@@
 
- 	\item	Шаг поиска будет успешным, если значения целевой функции в контрольных точках не превышают исходных значений целевой функции.
+ 	\item	Сделать шаг по текущей координате. Шаг поиска будет успешным, если значения целевой функции в контрольных точках не превышают исходных значений целевой функции.
 
@@
 
+ На рисунке (\ref{fig:fig1}) приведен небольшой пример работы алгоритма.
+Отображены линии уровней функции, закрашенными  кружками обозначены успешные шаги, не закрашенные --- неуспешные шаги при исследующем поиске.
+
  \begin{figure}[!h]
@@
 		\begin{minipage}[h]{0.8\linewidth}
-			\includegraphics[width=1\linewidth]{figure/fig1.png}
+			\includegraphics[width=1\linewidth]{figure/fig1.pdf}
 			\caption{Пример итераций алгоритма Хука-Дживса} %% подпись к рисунку
@@
 Дерево решений представляет собой бинарное дерево (дерево, в котором каждый не листовой узел имеет два дочерних узла).
-Такая структура позволяет использовать подобное дерево в машинном обучении (причем как для решения задач классификации, так и для регрессионных задач) в целях автоатизации анализа большого объема данных.
+Такая структура позволяет использовать подобное дерево в машинном обучении (причем как для решения задач классификации, так и для регрессионных задач) в целях автоматизации анализа большого объема данных.
 Для задач регрессии, каждому листу дерева присваивается константа, как следствие  получаемая функция аппроксимации является кусочно-постоянной.
@@
 
-В нашем случае дерево решений строит функцию $\varphi(y)$ кусочно постоянную аппроксимацию $\varphi(y)$  в многомерном пространстве, обозначим как $z' = \psi(y)$ значение вычисленное по decision tree в точке $y$.
+В нашем случае дерево решений строит функцию $\varphi(y)$ кусочно постоянную аппроксимацию $\varphi(y)$  в многомерном пространстве, обозначим как $z' = \psi(y)$ значение вычисленное по дереву решения в точке $y$.
 
-Библиотека алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения с открытым кодом OpenCV предоставляет уже готовые к использованию деревья решений.
+Библиотека алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения с открытым кодом OpenCV \cite{Brahmbhatt2013} предоставляет уже готовые к использованию деревья решений.
 В данной работе, при разработке алгоритма, применяющего деревья решений для выявления областей притяжения локальных минимумов, мы использовали данную библиотеку.
 И, следовательно, все подробности внутренней реализации деревьев решений можно найти в соответствующей документации - \cite{fio_bib16}
+
+\begin{figure}[ht!]
+	\begin{center}
+			\begin{minipage}[h]{0.8\linewidth}
+					\includegraphics[width=1\linewidth]{figure/Figure_1}
+					\caption{Функция Растригина} %% подпись к рисунку
+					\label{fig:fig2}
+				\end{minipage}
+		\end{center}
+\end{figure}
+
+\begin{figure}[ht!]
+	\begin{center}
+			\begin{minipage}[h]{0.8\linewidth}
+					\includegraphics[width=1\linewidth]{figure/Figure_2}
+					\caption{Аппроксимация по двумерному дереву решений} %% подпись к рисунку
+					\label{fig:fig2_2}
+				\end{minipage}
+		\end{center}
+\end{figure}
+
+На рисунке  (\ref{fig:fig2}) отображена трехмерная поверхность  оптимизационной задачи (Функции Расстрига), и на рисунке  (\ref{fig:fig2_2}) отображена аппроксимация   построенная по дереву решений.
 
@@
 Точки, расположенные близко в многомерном пространстве, могут оказаться на разных концах одномерного отрезка и наоборот).
-
-%\begin{figure}[ht!]
-%	\begin{center}
-	%		\begin{minipage}[h]{0.8\linewidth}
-		%			\includegraphics[width=1\linewidth]{figure/fig2.png}
-		%			\caption{Функция из генератора gkls до использования кривых Пеано} %% подпись к рисунку
-		%			\label{fig:fig2}
-		%		\end{minipage}
-	%	\end{center}
-%\end{figure}
-
-%\begin{figure}[ht!]
-%	\begin{center}
-	%		\begin{minipage}[h]{0.8\linewidth}
-		%			\includegraphics[width=1\linewidth]{figure/fig2.png}
-		%			\caption{Рисунок с двумерным деревом решений} %% подпись к рисунку
-		%			\label{fig:fig2_2}
-		%		\end{minipage}
-	%	\end{center}
-%\end{figure}
 
@@
 	\item  Перенумеровать (нижним индексом) точки ранее проведенных испытаний $x^i, 1\leq i\leq k$, а также граничные точки отрезка [0,1] в порядке возрастания координаты:
-	\begin{equation}
-	\label{agp1_sort}
+	\begin{displaymath}	
 	0=x_0<\ x_1<\ ...\ <x_{k+1}=1.
-	\end{equation}
+	\end{displaymath}
 	и сопоставить им значения $z_i=f(x_i)$.
@@
 	\item  Вычислить текущие нижние оценки $M$ неизвестной константы Гельдера $H$:
-	\begin{equation}
-	\label{agp2_mu}
+	\begin{displaymath}	
 	\mu=max\left\{\frac{|z_i-z_{i-1}|}{{{(x}_i-x_{i-1})}^{1/N}},\ i=1,\ldots,k\right\},\ M=\ \left\{\begin{matrix}r\mu,\ \mu>0,\\1,\ \mu=0,\\\end{matrix}\right.\
-	\end{equation}
+	\end{displaymath}
 	где $r>1$ --- параметр алгоритма.
@@
 	\item  Для всех интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить значение $R(i)$, называемое \textit{характеристикой} интервала, в соответствии с формулами
-	\begin{equation}
-	\label{agp3_R1}
+	\begin{displaymath}
 	R(1)=2\Delta_1-4\dfrac{z_1}{M}, \; R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M},
-	\end{equation}
-	\begin{equation}
-	\label{agp3_Ri}
+	\end{displaymath}
+	\begin{displaymath}
 	R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1,
-	\end{equation}
+	\end{displaymath}
 	где \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
@@
 	\item   Упорядочить характеристики $R\left(i\right),\ 1\leq i \leq k+1,$ в порядке невозрастания 
-	\begin{equation}
-	\label{agp4_R_sort}
+	\begin{displaymath}
 	R\left(t_1\right)\geq\ R\left(t_2\right)\geq...\geq\ R\left(t_k\right)\geq\ R(t_{k+1}),\ 
-	\end{equation}	
+	\end{displaymath}	
 	и выбрать интервал с номерами $t$, с наибольшими значениями характеристики.
@@
 	\item  В выбранном интервале вычислить точку $x^{k+1}$, в соответствии с формулами
-	\begin{equation}
-	\label{agp5_x1}
+	\begin{displaymath}
 	x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\ t_j=1,\ t_j=k+1,
-	\end{equation}	
-	\begin{equation}
-	\label{agp4_xi}	
+	\end{displaymath}	
+	\begin{displaymath}
 	x^{k+1}=\frac{x_{t_j}+x_{t_j-1}}{2}-sign\left(z_{t_j}-z_{t_j-1}\right)\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu}\right]^N,\ 1<t_j<k+1.
-	\end{equation}
+	\end{displaymath}
 
@@
 
-	\item Создаем decision tree по множеству $I_k$, получаем функцию аппроксимации  $\psi(y)$.
+	\item Создаем дерево решений по множеству $I_k$, получаем функцию аппроксимации  $\psi(y)$.
 
-	\item 	Если используем decision tree впервые, то строим равномерную сетку
+	\item 	Если используем дерево решений впервые, то строим равномерную сетку
 
@@
 
-%На рисунке  (\ref{fig:fig2}) отображены линии уровней оптимизационной задачи с отображением текущих  точек испытаний, и на рисунке  (\ref{fig:fig2_2}) отображено соответствующее дерево решений.
-
 Все полученные результаты были добавлены в систему globalizer \cite{fio_bib18}.
@@
 Использовалась данная система при реализации схемы с запуском локальных методов в нужный момент.
-Ниже приведена блок-схема того, как данная схема была объединена с АГП.
+Ниже приведена блок-схема того, как данная схема была объединена с АГП (\ref{fig:fig3}).
 

### 2023-02-01 2023/PAVT/ParallelDecisionTree.tex
Исправлены небольшие баги. Добавлена ссылка на рисунок.

@@
 
+ Fig.
+(\ref{fig:fig1}) presents small example of how the algorithm works.
+Function contour plot are displayed, filled circles indicate successful steps, unfilled circles indicate unsuccessful steps in the investigating search.
+
 \subsection{Decision tree}\label{SecDT}
@@
 				\begin{minipage}[h]{0.9\linewidth} 			\includegraphics[width=1\linewidth]{figure/fig3.pdf}
-					 			\caption{Block scheme of combined GSA and Hooke-Jeeves method using the  decision tree} %% подпись к рисунку 			\label{fig:fig3}
+					 			\caption{Block scheme of combined GSA and Hooke-Jeeves method using the  decision tree} %% подпись к рисунку 	
+					 			\label{fig:fig3}
 				 \end{minipage}
@@
 Table \ref{tab:1} presents the averaged numbers of iterations executed by respective methods when  solving a series of GKLS problems.
-The sign ``>'' points to the situation when not all problems are  solved.
+The sign `` $>$ '' points to the situation when not all problems are  solved.
 In this case, the number of non-solved problems is given in brackets.
@@
         N & Problem class & DIRECT & DIRECTl & GSA  \\ \hline
-        4 & Simple & >47282 (4) & 18983 & 11953  \\ \hline
-        ~ & Hard & >95708 (7) & 68754 & 25263  \\ \hline
-        5 & Simple & >16057 (1) & 16758 & 15920  \\ \hline
-        ~ & Hard & >217215 (16) & >269064 (4) & >148342 (4)  \\ \hline
+        4 & Simple & $>47282 (4)$ & 18983 & 11953  \\ \hline
+        ~ & Hard & $>95708 (7)$ & 68754 & 25263  \\ \hline
+        5 & Simple & $>16057 (1)$ & 16758 & 15920  \\ \hline
+        ~ & Hard & $>217215 (16)$ & $>269064 (4)$ & $>148342 (4)$  \\ \hline
     \end{tabular}

### 2023-02-01 2023/PAVT/ParallelDecisionTree.tex
minor corrections

--- 2023-01-31 2023/PAVT/ParallelDecisionTree.tex
@@
 
-The investigating search is performed as follows. 
+\begin{figure} 
+	\begin{center} 
+		\begin{minipage}[h]{0.8\linewidth} 
+			\includegraphics[width=1\linewidth]{figure/fig1.pdf} 
+			\caption{An example of iterations of the Hooke-Jeeves algorithm} %%  подпись к рисунку 
+			\label{fig:fig1} 
+		\end{minipage} 
+	\end{center} 
+\end{figure}
+
+The investigating search is performed as follows (see Fig. \ref{fig:fig1}).
 \begin{enumerate} 
@@
 The magnitude of this step is defined by a parameter set in advance.
-
-\begin{figure}[!h] 
-	\begin{center} 
-		\begin{minipage}[h]{0.8\linewidth} 
-			\includegraphics[width=1\linewidth]{figure/fig1.pdf} 
-			\caption{An example of iterations of the Hooke-Jeeves algorithm} %%  подпись к рисунку 
-			\label{fig:fig1} 
-		\end{minipage} 
-	\end{center} 
-\end{figure}
 
@@
 Let us denote the value  computed by the  decision tree at the point $y$ as $z' = \psi(y)$.
+Fig. \ref{fig:fig2} presents a graph of the objective function $\varphi(y)$; Fig. \ref{fig:fig2_2} presents corresponding approximation $\psi(y)$ built using decision tree.
+
+\begin{figure}
+	\begin{center}
+		\begin{minipage}[h]{0.8\linewidth}
+			\includegraphics[width=1\linewidth]{figure/fig5.png}
+			\caption{An objective function $\varphi(y)$ from the test class} %% подпись к рисунку
+			\label{fig:fig2}
+		\end{minipage}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+		\begin{minipage}[h]{0.8\linewidth}
+			\includegraphics[width=1\linewidth]{figure/fig4.png}
+			\caption{A piecewise-constant  approximation $\psi(y)$ of the function $\varphi(y)$} %% подпись к рисунку
+			\label{fig:fig2_2}
+		\end{minipage}
+	\end{center}
+\end{figure}
 
@@
 Only in the case when  all function values at the neighboring points are greater than at the current point, we can run the local  method.
-
-Fig. \ref{fig:fig2} presents a contour plot of an objective function and the trial points; Fig.  \ref{fig:fig2_2} presents corresponding decision tree.
-
-\begin{figure}[ht!]
-	\begin{center}
-		\begin{minipage}[h]{0.9\linewidth}
-			\includegraphics[width=1\linewidth]{figure/fig4.png}
-			\caption{A function from the GKLS generator before using Peano curves} %% подпись к рисунку
-			\label{fig:fig2}
-		\end{minipage}
-	\end{center}
-\end{figure}
-
-\begin{figure}[ht!]
-	\begin{center}
-		\begin{minipage}[h]{0.9\linewidth}
-			\includegraphics[width=1\linewidth]{figure/fig5.png}
-			\caption{A figure with a two-dimensional decision tree} %% подпись к рисунку
-			\label{fig:fig2_2}
-		\end{minipage}
-	\end{center}
-\end{figure}
 

### 2023-02-01 2023/PAVT/ParallelDecisionTree.tex

--- 2023-02-01 2023/PAVT/ParallelDecisionTree.tex
@@
 %
-\title{On the use of decision trees to identify the local extrema in parallel global optimization algorithms}
+\title{On using the decision trees to identify the local extrema in parallel global optimization algorithms}
 %
-\titlerunning{Global Optimization ...}  % abbreviated title (for running head)
+\titlerunning{On using the decision trees to identify the local extrema }  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 %
-\author{K.~A.~Barkalov\inst{1}\letter\orcidID{0000-0001-5273-2471} \and I.~G.~Lebedev\inst{1}\letter\orcidID{0000-0002-8736-0652} \and D.~I.~Silenko\inst{1}\letter\orcidID{0000-0002-2578-9699}}
+\author{Konstantin Barkalov\inst{1}\letter\orcidID{0000-0001-5273-2471} \and Ilya Lebedev\inst{1}\letter\orcidID{0000-0002-8736-0652} \and Dmitry Silenko\inst{1}\letter\orcidID{0000-0002-2578-9699}}
 %
-\authorrunning{K.~A.~Barkalov et al.} % abbreviated author list (for running head)
+\authorrunning{Konstantin Barkalov et al.} % abbreviated author list (for running head)
 %
 %%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{K.~A.~Barkalov and I.~G.~Lebedev, and D.~I.~Silenko}
+\tocauthor{K.A.~Barkalov and I.G.~Lebedev, and D.I.~Silenko}
 %
@@
 
-The investigating search is performed as follows. 
-\begin{enumerate} 
-\item	The step value is determined (it is different for each coordinate and can be varied in the course  of search).  
-\item	The search step is considered to be successful if the value of the objective function at the check  point doesn't exceed the value of the objective function at the initial point.  
-\item	Otherwise, it is necessary to return to the previous item and make a step in the reverse direction.  
-\item	After making the steps in all $N$ coordinates, the investigating search is completed. The point  obtained is called the base point. 
-\end{enumerate}
-
-With regard to the search in a direction, it consists in performing a step from the base point found  during the investigating search along the straight line direction connecting this one with the previous  base point.
-The magnitude of this step is defined by a parameter set in advance.
-
-\begin{figure}[!h] 
+\begin{figure} 
 	\begin{center} 
@@
 
+The investigating search is performed as follows.
+\begin{enumerate} 
+\item	The step value is determined (it is different for each coordinate and can be varied in the course  of search).  
+\item	The search step is considered to be successful if the value of the objective function at the check  point doesn't exceed the value of the objective function at the initial point.  
+\item	Otherwise, it is necessary to return to the previous item and make a step in the reverse direction.  
+\item	After making the steps in all $N$ coordinates, the investigating search is completed. The point  obtained is called the base point. 
+\end{enumerate}
+
 Fig.
-(\ref{fig:fig1}) presents small example of how the algorithm works.
-Function contour plot are displayed, filled circles indicate successful steps, unfilled circles indicate unsuccessful steps in the investigating search.
+\ref{fig:fig1} shows the example of the algorithm work.
+Function level lines are displayed, filled circles indicate successful steps, unfilled circles correspond to unsuccessful steps performed during the investigating search.
+With regard to the search in a selected direction, it consists in performing a step from the base point found during the investigating search along the line connecting this one with the previous base point.
+The magnitude of this step is defined by a parameter set in advance.
 
@@
 The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard from \cite{Sergeyev2006}) of dimensionalities 2, 3, 4, and 5.
+These two classes differ in the size of the attraction region of the global minimum point. 
+For the simple class of problems, the radius of the attraction region is three times larger than for a complex one.
+
 The stop  criterion for the algorithms was the falling of the next trial points into the $\varepsilon$-nearness of  true global minimum.
@@
 As one can see from Table \ref{tab:2},  the iteration speedup is large enough for the problems of any  dimensionality. 
-
 So far, as a result of the work done, we succeeded to combine the global search algorithm with a local  optimization method.
@@
 The proposed scheme allows us utilizing the advantages of both parallelization as well as fast search of local extrema.
+
+In addition to using decision trees to identify attraction regions for local extrema of multiextremal functions, we also plan to use machine learning methods to separate the variables of the problem being solved.
+In many applied optimization problems the dependence of the objective function on some parameters is either linear or unimodal.
+Separating such variables into a special group and solving the problem using a parallel recursive optimization scheme \cite{Barkalov2020_1} can reduce the time for solving the problem by orders of magnitude compared to using a global search for all variables at once.
+
+\medskip
+
+\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204 .
 

### 2023-02-01 2023/PAVT/ParallelDecisionTree.tex
minor corrections before submitting

@@
 
-\section{The algorithms used}\label{SecA}
+\section{Algorithms}\label{SecA}
 
@@
 
-The algorithm stops in the case if the condition \(\Delta_{t_j} < \varepsilon\) is satisfied for even a  single value $t_j,\ 1\le\ j\le\ p$ from (\ref{agp4_R_sort}).
-This stop criterion (along with the criterion limiting the number of executed iterations usual for the  iteration methods) is used in the applied optimization problems, in which the global minimum point $x^*$ is unknown a priori.
+The algorithm stops in the case if the condition \(\Delta_{t_j} < \varepsilon\) is satisfied for even a  single value $t_j,\ 1\le\ j\le\ p$, from (\ref{agp4_R_sort}).
+This stop criterion (along with the criterion limiting the number of executed iterations usual for the  iteration methods) is used in the applied optimization problems, in which the global minimum point $y^*$ is unknown a priori.
 
@@
 
-After completing a certain number of trials (for example, $100\ \ast\ N$), let us use all accumulated  points to initialize a suitable data structure and training the decision tree on its base.
+After completing a certain number of trials (for example, $100\cdot N$), let us use all accumulated  points to initialize a suitable data structure and training the decision tree on its base.
 To make the  determining the adjacent points easier, we construct a uniform grid with certain step in each coordinate.
@@
 	\item If the decision tree is used for the first time, construct a uniform grid
-
 	\begin{displaymath} 
-		Y'=\{ y'\in\ R^N:\ a_i\le\  {y'_i}^k \le\ b_i,\ 1\le\ i\le\ N,\ 1\le k\le\ \sqrt[N]{300}  \} 
+		Y'=\{ y'\in R^N:\ a_i\le y'_i \le b_i,\ 1\le i\le N \},
 	\end{displaymath} 
+	where number of the grid nodes is a parameter of the method.
 

### 2023-02-01 2023/PAVT_Ufa/PAVT2023.tex
the version submitted

@@
-%%%%%%%%%%%%%%%%%%%% author.tex 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-%%%%%%%%%%%%%%%% Springer 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %
-% RECOMMENDED 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%
+% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 \usepackage{marvosym}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{cite}
-
-\usepackage[russian]{babel}
 
@@
 \mainmatter              % start of a contribution
-% Решение обратных задач химической кинетки с помощью асинхронного алгоритма глобального поиска
-% Решение обратных задач химической кинетки с помощью смешанного локально-глобального поискового алгоритма 
-%\title{Solving the Inverse Problems of Chemical Kinetics Using the Asynchronous Global Optimization Algorithm}
-\title{
-Solving ...
-}
-\titlerunning{Solving the Inverse Problems of Chemical Kinetics}  % abbreviated title (for running head)
+%
+\title{Parallel Computing in Tikhonov Regularization Method for Solving the Inverse Problem of Chemical Kinetics}
+
+%Parallel computing in Tikhonov regularization method for solving the inverse problem of chemical kinetics
+
+%Solving the Regularized Inverse Problem of Chemical Kinetics Using the Parallel Optimization Algorithm
+
+%
+\titlerunning{Parallel Computing}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 %
-\author{
-Irek Gubaydullin$^{1,2}$\and
-Leniza Enikeeva$^{2,3}$\orcidID{0000-0003-4219-4870}
-\and
-Konstantin Barkalov$^4$ \letter \orcidID{0000-0001-5273-2471}
-\and
-Marina Usova$^4$\orcidID{XXX} 
-}
+\author{Konstantin Barkalov\inst{1}\letter\orcidID{0000-0001-5273-2471} \and Marina Usova\inst{1}\orcidID{0000-0002-0722-6884} \and Leniza Enikeeva\inst{2}\letter\orcidID{0000-0003-4219-4870} \and Dmitry Dubovsev\inst{3} \and Irek Gubaydullin\inst{2,3}}
 
 %
-\authorrunning{I. Gubaydullin et al.} % abbreviated author list (for running head)
+\authorrunning{K. Barkalov et al.} % abbreviated author list (for running head)
 %
 %%%% list of authors for the TOC (use if author list has to be modified)
-%\tocauthor{Konstantin Barkalov and Ilya Lebedev }
+%\tocauthor{Ivar Ekeland and Kim~B.~Bruce, and Roger Temam}
 %
-
-\institute{$^1$Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia\\$^2$Ufa State Petroleum Technological University, Ufa, Russia \\$^3$Novosibirsk State University, Novosibirsk, Russia\\$^4$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
-\email{leniza.enikeeva@yandex.ru},
-\email{konstantin.barkalov@itmm.unn.ru},
-\email{ilya.lebedev@itmm.unn.ru}
+\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
+	\email{konstantin.barkalov@itmm.unn.ru}
+	\and
+	Ufa State Petroleum Technological University, Ufa, Russia \\
+	\email{leniza.enikeeva@yandex.ru} \\
+        Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia
 }
@@
 \begin{abstract}
+% Проверьте первое предложение - мой вариант ниже
+% The authors investigates the advantages and disadvantages of the products refining processes used in the gasoline production. 
+The authors investigates the advantages and disadvantages of the products refining processes used in the gasoline production.
+As one of the ways to solve the problem of environmental friendliness of motor fuel, it is proposed to consider increasing the practice of involving alkylate in compounding and improving its quality.
+The work describes the scheme of chemical transformations of the sulfuric acid alkylation process, taking into account the target reactions and side effects.
+Based on the studied chemistry of the reactions, the authors pose the inverse problem of chemical kinetics and consider the regularization method for its solution.
+The global optimization problem, which corresponds to the regularized inverse problem, was solved using a parallel optimization algorithm.
+The results of a computational experiment on a supercomputer are provided, showing the adequacy of the solution obtained.
 
-The paper considers ...
-%the application of parallel computing technology to the simulation of catalytic chemical reaction, which is widely used in the modern automobile industry to produce gasoline with high octane number.
-As a chemical reaction, the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid is assumed.
-To simulate a chemical process, it is necessary to develop a kinetic model of the process, that is, to determine the kinetic parameters.
-To do this, the inverse problem of chemical kinetics is solved, which predicts the values of kinetic parameters based on laboratory data.
-From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem.
-A parallel information-statistical global search algorithm was used to solve it.
-The use of the parallel algorithm has significantly reduced the search time to find the optimum.
-The found optimal parameters of the model made it possible to adequately simulate the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid.
+% We would like to encourage you to list your keywords within
+% the abstract section using the \keywords{...} command.
+\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Chemical kinetics $\cdot$ Inverse problems $\cdot$ Regularization}
+\end{abstract}
+%
+\section{Introduction}
+%
+Currently, in order to meet the requirements of environmental protection, the standards of clean gasoline in the world are developing in the directions of low sulfur content, olefins, aromatic substances and high octane.
+This means that the refining industry requires stricter fuel product standards and cleaner production processes.
+Gasoline is obtained during the compounding of products of several refining processes: catalytic reforming, isomerization, catalytic cracking and the process of sulfuric acid alkylation.
+The use of each component is limited by certain factors, including the sulfur content, saturated vapor pressure, the content of aromatic hydrocarbons and the octane number of the final product \cite{alkylation_mod}.
 
-%обновить ключевые слова
-\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Chemical kinetics $\cdot$ Inverse problems }
-\end{abstract}
+The product of catalytic cracking is a catalyst obtained as a result of cracking of vacuum gas oil (fraction 350 - 500 $^0$C).
+It contains aromatic hydrocarbons, olefins, and a small amount of hydrocarbons from construction.
+It is a source of sulfur in commercial gasoline.
 
-\section{Introduction}
+In the process of catalytic isomerization of light gasoline fractions (i.b.p-62 $^0$C), a high-octane component isomerizate is obtained.
+The octane number of isomerizate reaches up to 92 points according to the research method, however, its excessive involvement during compounding leads to an increase in the pressure of saturated vapors in commercial gasoline and the formation of steam plugs in the power system during the hot season\cite{alkylation_mod}.
 
-%Часть УГНТУ
-%Currently, there is a tendency to improve the environmental characteristics of automobile fuel while maintaining a high octane number.
-Sulfuric acid alkylation of isobutane with olefins makes it possible to obtain a high-octane component of gasoline with a minimum content of aromatic hydrocarbons.
-The alkylate, which is produced by the alkylation of isobutane with C3-C5 olefins in the presence of strong acid, has the advantages of high octane number, low vapor pressure, and zero content of olefins and aromatics that allow it to be a desirable blending component for high-quality gasoline.
-Alkylates will continue to act as a desirable blending component for high-quality gasoline as the quality of gasoline continues to increase [ссылка на китайскую статью].
-Therefore, it is a significant process of the modern refinery [ссылка на catalyst].
-To optimize the chemical process in industry, it is necessary to first develop its model, which in this case means building a mathematical model of the chemical process and then its kinetic model, that is, numerically calculate the kinetic constants of the reaction.
+The source of aromatic contents in gasoline is reformat, obtained as a result of catalytic reforming of heavy gasoline fractions (fraction 85 (105)-180 $^0$C).
+The involvement of lighter fractions in the reforming raw materials leads to an increase in the proportion of benzene in the reformate.
+
+Alkylated gasoline fully complies with the operational and environmental requirements of modern European, American and Russian industrial standards for the production of fuels for automotive internal combustion engines and is an ideal and essential component of reformed environmentally friendly gasoline.
+The volume of alkylate production abroad exceeds 70 million tons per year, and in the Russian Federation in 2019 reached the level of 2 million tons per year.
+Alkylate is obtained as a result of alkylation of isoalkanes by alkenes in the presence of a catalyst.
+The most common process catalyst in Russia is sulfuric acid.
+The advantages of alkylate include a high octane number (up to 90 points according to the research method), low saturated vapor pressure, low content of heteroatomic compounds and good chemical stability.
+In addition, the sensitivity of alkylate does not exceed 5 points \cite{cao2019}.
+
+There are target and side reactions of the sulfuric acid alkylation process, which proceeds by the carbonium-ion mechanism. The process is carried out in several stages:
+
+1. The first stage is the addition of an acid proton to an olefin to produce a tert-butyl carbation:
+\begin{figure}
+\centering
+\includegraphics[]{fig1.png}
+\caption{The addition of an acid proton to an olefin to produce a tert-butyl carbation}
+\label{fig1}
+\end{figure}
+
+2.
+In the second stage, the formed carbonium ion interacts with paraffin hydrocarbon.
+In this case, the hydrogen anion from the tertiary carbon atom of the isoparaffin hydrocarbon passes to the carbonium ion formed at the first stage of the reaction:
+
+\begin{figure}
+\centering
+\includegraphics[]{fig2.png}
+\caption{The interaction of the formed carbonium ion with paraffin hydrocarbon}
+\label{fig2}
+\end{figure}
+
+3. The third stage consists in the addition of a tertiary carbonium ion to the second olefin molecule to form carbonium:
+
+\begin{figure}
+\centering
+\includegraphics[]{fig3.png}
+\caption{The addition of the tertiary carbonium ion to the second olefin molecule}
+\label{fig3}
+\end{figure}
+
+4. The fourth stage consists in the skeletal isomerization of the secondary carbonium ion:
+
+\begin{figure}
+\centering
+\includegraphics[height=4.3 cm]{fig4.png}
+\caption{The skeletal isomerization of secondary carbonium ion}
+\label{fig4}
+\end{figure}
+
+5.
+The fifth stage is the interaction of the formed carbonium ions with an isoparaffin molecule by a tertiary carbon-hydrogen bond with the formation of target products and a carbonium ion from isoparaffin:
+
+\begin{figure}
+\centering
+\includegraphics[height=3.5 cm]{fig5.png}
+\caption{The interaction of the formed carbonium ions with the isoparaffin molecule}
+\label{fig5}
+\end{figure}
+
+For ease of use when compiling the model, we will make up a chain of transformations of the sulfuric acid alkylation process (Figure 6).
+
+\begin{figure}
+\centering
+\includegraphics[height=6.5 cm]{fig6.png}
+\centering
+\caption{The scheme of reactions of the sulfuric acid alkylation process}
+\label{fig6}
+\end{figure}
+
+It should be noted that the temperature range of the process is from 2 to 15 degrees.
+Determining the optimal value for a given composition of raw materials relates to the objectives of the study.
+In this regard, the solution of the problem of constructing a model of the process of sulfuric acid alkylation of isobutane with butylenes is relevant\cite{alkylation_mod}.
+
 %Стыковочный абзац
-
-Since the mathematical model of the chemical reaction is a system of differential equations, it is only possible to find the values of the constants in this system numerically (see, e.g., \cite{RSD}).
+Since the mathematical model of the chemical reaction is a system of differential equations, it is only possible to find the values of the constants in this system numerically (see, e.g., \cite{Gubaydullin2021}).
 Note that the objective function in such problems is usually multiextremal, i.e.
 it has many local extrema along with a global one.
+
 %Часть ННГУ
-Numerical methods for solving such multiextremal problems (global optimization methods) differ significantly from local search methods (see, e.g., \cite{Sergeyev2017, PaulaviciusZilinskas2014}).
+%Numerical methods for solving such multiextremal problems (global optimization methods) differ significantly from local search methods (see, e.g., \cite{PaulaviciusZilinskas2014,Sergeyev2017}).
 From an algorithmic point of view, they can be divided into two classes: metaheuristic and deterministic.
@@
 For example, we can assume that the ratio of the increment of the function to the corresponding increment of the argument cannot exceed some threshold.
-In this case, the functions are called Lipschitz functions, and the problem itself is a Lipschitz global optimization problem.
+In this case, the functions are called Lipschitzian, and the problem itself is a Lipschitz optimization problem.
 This paper continues the series of papers where the parallel methods of Lipschitz optimization proposed in \cite{Strongin2000} are investigated and modified in their application to solving inverse problems of chemical kinetics.
@@
 Section \ref{Sec_math_mod} describes the mathematical model of the chemical reaction under study.
-The formal statement of the Lipschitz global optimization problem and the general scheme of search algorithms are given in Section \ref{Sec_GO_Problem}.
-In Section \ref{Sec_PGSA}, a scheme of the proposed asynchronous parallel algorithm for solving multiextremal problems is presented.
+The formal statement of the Lipschitz global optimization problem and the general scheme of search algorithms are given in Section \ref{Sec_GSA}.
+In this section a scheme of the proposed asynchronous parallel algorithm for solving multiextremal problems is also presented.
 The results of numerical solution of the inverse problem of chemical kinetics are discussed in Section \ref{Sec_Exp}.
 
-\section{Problem Statement}\label{Sec_math_mod}
-%Содержательная постановка задачи
+\section{Problem Statement} \label{Sec_math_mod}
 
-\section{Parallel Algorithm for Solving Global Optimization Problems }\label{Sec_GSA}
+To optimize the technological process, it is necessary to know the kinetic laws and mechanism of chemical reactions, to determine which it is necessary to evaluate the kinetic parameters of chemical reactions by solving the inverse problem of chemical kinetics\cite{regularization}.
+In most cases, the equations of chemical kinetics are systems of ordinary nonlinear differential equations for concentrations of substances $x_{i}$ (\ref{concentration_changing}) with initial conditions (\ref{init_cond}).
 
-\subsection{Global Optimization Problem}\label{Sec_GO_Problem}
+\begin{equation}
+  \frac{dx_{i}}{dt} = \sum_{j=1} ^J \nu_{ij} \omega_j,i=1,....I,\omega_j=k_j^0 \cdot \exp (-\frac{E_{j}}{RT})\cdot \prod_{i=1} ^M x_{i}^{|a_{ij}|}
+\label{concentration_changing}
+\end{equation}
 
-As it was mentioned above, the problem of identifying mathematical model parameter values can be considered a Lipschitz global optimization problem.
+\begin{equation}
+t=0, x_i(0)=x_i^0,
+\label{init_cond}
+\end{equation}
+
+moreover, $k_j^0 \cdot \exp (-\frac{E_{j}}{RT})$ they are called the rate constants of the reaction stages, where $I$ is the number of substances involved in the reaction; $J$ is the number of reaction stages; $v_{ij}$ are stoichiometric coefficients; $E_j$ are the activation energies of reactions, cal/mol; $R$ is the universal gas constant, cal /(mol $\cdot$ K); $T$ – temperature, K; $a_{ij}$ – stoichiometric coefficients $k_j^0$ – pre-exponential multipliers.
+
+The inverse problem of chemical kinetics is a global optimization problem, implying the need to determine the vector of reaction rate constants ($k_1, k_2, ..., k_J$), at which the deviation of the calculated concentrations of the reaction components from the experimental ones will be minimal.
+Thus, in order to determine the rate constants of the reaction, it is necessary to solve the inverse problem of chemical kinetics by repeatedly solving the direct problem – by iterating over the rate constants of the stages (or a set of pre-exponents and activation energies) according to some algorithm.
+To solve the optimization problem, it is necessary to minimize the functional  (\ref{functional}) deviation of experimental data from the calculated ones:
+
+\begin{equation}
+FF = \sum_{i=1} ^M \sum_{j=1} ^N |x_{ij}^{calc} - x_{ij}^{exp}|,
+\label{functional}
+\end{equation}
+where $x_{ij}^{calc}$ are calculated values; $x_{ij}^{exp}$ are experimental data; $M$ is the number of experimental points; $N$ is the number of substances involved in the reaction.
+To solve the instability problem of optimization problem (\ref{functional}), consider the minimization problem in general form
+
+\begin{equation}
+f(z) \rightarrow \inf, z \in D,
+\label{problem}
+\end{equation}
+where $D$ is some given set, and $f: D \rightarrow R^1$ is the function (functional) specified on it.
+
+In this inverse problem of chemical kinetics, a number of kinetic parameters act as a point $z$ (as a rule, a vector of reaction rate constants), while the function $f$ is the deviation of the calculated concentrations of reaction components from empirical data (functional (\ref{functional})).
+Problem (\ref{problem}) can be attributed to one of two classes of problems, which in the literature are called correctly posed and incorrectly posed minimization tasks, respectively.
+Inverse problems of chemical kinetics are incorrectly posed problems.
+According to Hadamard, a problem is considered to be posed correctly if 1) its solution exists, 2) the solution is unique, 3) the solution is stable to variations in the initial data.
+If at least one of the listed requirements is not met, the task is considered incorrectly set.
+The inverse problems of chemical kinetics encountered in practice usually have a solution and condition 1 is thus met.
+However, conditions 2 and 3 are most often not met\cite{regularization}.
+For the approximate solution of ill-posed problems, special methods are required that make it possible to reliably construct a solution to problem (\ref{problem}).
+Such methods exist and are commonly called regularization methods.
+In this paper, the regularization method of A.N.
+Tikhonov, or the stabilization method, is used, and the classical machine learning apparatus is used – normalization of features and scaling.
+The main task that we will consider will be the task of minimizing
+
+\begin{equation}
+f^0(z) \rightarrow \inf, z \in D \subset Z,
+\label{minimization_problem}
+\end{equation}
+in which we consider the set $D$ to be a convex and closed set of Hilbert space $Z$.
+The function $f^0:D \rightarrow R^1$ will be considered continuous and convex on $D$.
+Suppose that problem (\ref{minimization_problem}) has a solution, i.e.
+the set $D^*$ is nonempty and denote by $z^0$ the norm-minimal solution of this problem.
+Suppose that instead of an exact function $f$, we know its approximation $f_{\delta},$ $\delta \in [0, \delta_0], \delta_0>0$ which is some sufficiently small number, and
+
+\begin{equation}
+|f^\delta(z) - f^0(z)| \leq \delta (1+||z||^2),  \forall z \in D. 
+\end{equation}
+
+The main construction of the Tikhonov method is the smoothing function or Tikhonov function
+
+\begin{equation}\label{formula_T}
+T_\alpha^\delta(z) \equiv f^\delta(z) + \alpha ||z||^2, z \in D.
+\end{equation}
+where $\alpha > 0$ is the regularization parameter. The term $\alpha||z||^2$ is called the stabilizing term. Consider the auxiliary minimization problem
+
+\begin{equation}
+T_\alpha^\delta(z) \rightarrow \inf, z \in D. 
+\label{tikhonov_F}
+\end{equation}
+
+Various numerical methods can be used to approximate the solution of this problem. Suppose that as a result of a finite number of iterations we get a point $z_\alpha^{\delta,\epsilon}$ such that
+
+\begin{equation}
+T_\alpha^\delta(z) \equiv \min T_{\alpha}^{\delta}(z) \leq T_{\alpha}^{\delta} (z_{\alpha}^{\delta,\varepsilon}) \leq T_{\alpha}^{\delta} + \varepsilon,
+\end{equation}
+where the value $\varepsilon > 0$ characterizes the accuracy of solving the minimization problem (\ref{tikhonov_F}).In this case, the minimization problem (\ref{tikhonov_F}) for each fixed pair $\delta, \alpha$ has, as a rule, a much larger ``margin of stability'' than the original problem (\ref{problem}) and in most cases is correctly posed.
+
+\section{Parallel Global Optimization Algorithm}\label{Sec_GSA}
+
+\subsection{Optimization Problem}\label{Sec_GO_Problem}
+
+As it was mentioned above, the problem of identifying the model parameter values can be considered a Lipschitz global optimization problem.
 From the formal point of view this problem is a mathematical programming problem of the form
-\begin{gather}
- \varphi^* = \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problemN}\\
- D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\} \label{D},
-\end{gather}
-where $a,b\in R^N$ are the given vectors and $\varphi(y)$ is the objective function which satisfies the Lipschitz condition
+\begin{equation} \label{problemN}
+ \varphi^* = \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},
+ \end{equation}
+\begin{equation} \label{D}
+ D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\},
+ \end{equation}
+where the vectors $a,b\in R^N$ correspond to the lower and upper bounds of the search region and $\varphi(y)$ is the objective function which satisfies the Lipschitz condition
 \begin{equation}\label{Lip}
@@
 
-The function $\varphi(y)$ is assumed to be multiextremal and is given as a ``black box'' (i.e.
+We suppose, that the function $\varphi(y)$ is multiextremal and is given as a ``black box'' (i.e.
 as some subroutine with a vector of parameters as its input and the calculated value of the function as its output).
@@
 There are several algorithms that can be applied to solve Lipschitz optimization problems.
-These include, among others, the non-uniform space covering method \cite{Evtushenko2013, Evtushenko2009}, diagonal and simplicial partitions methods \cite{Zilinskas2010, Paulavicius2011}.
-In this paper, we used the global search algorithm proposed by Strongin\cite{Strongin2000}.
+These include, among others, the non-uniform covering method \cite{Evtushenko2009,Evtushenko2013}, diagonal and simplicial partitions methods \cite{Zilinskas2010,Paulavicius2011}.
+In this paper, we used the global search algorithm proposed by Strongin \cite{Strongin2000}.
 Under this approach, the original multidimensional problem (\ref{problemN}) is reduced to a one-dimensional optimization problem using Peano-Hilbert curves.
 
-In fact, using a continuous one-to-one mapping (Peano-Hilbert curve) $y(x)$ of the segment $[0,1]$ of the real axis onto the hyperinterval $D$ from (\ref{D}), we can reduce the multidimensional problem (\ref{problemN}) to a one-dimensional problem
+In fact, using a continuous one-to-one mapping (Peano curve) $y(x)$ of the segment $[0,1]$ of the real axis onto the hyperinterval $D$ from (\ref{D}), we can reduce the original problem (\ref{problemN}) to a one-dimensional problem
 \[
@@
 \]
-where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
+where the one-dimensional function $\varphi(y(x))$ will satisfy a H{\"o}lder condition
 \[
@@
 with $ H=2 L \sqrt{N+3}$.
-The numerical construction of various approximations of the Peano-Hilbert curve is discussed in \cite{Strongin2000,Sergeyev2013}.
+The numerical construction of various approximations of such mappings is discussed in \cite{Sergeyev2013,Strongin2000}.
 
-Thus, the search trial at some point $x'\in[0,1]$ will involve first the construction of the image $y'=y(x')$, and only then the computation of the function value $ z' = \varphi(y')$.
+Thus, the search trial at some point $x'\in[0,1]$ will involve first the image construction $y'=y(x')$, and only then the function value computation $ z' = \varphi(y')$.
 
@@
 	\item 
-Calculate the function value $z^{k+1} = f(x^{k+1})$
+Calculate the function value $z^{k+1} = f(x^{k+1})$.
 
 	\item 
-Evaluate a global minimum estimate
+Evaluate a global minimum estimate.
 
@@
 To construct a concrete form of the interval characteristic, we can use a lower envelope (or minorant) of the function to be minimized or a mathematical expectation of function values, etc. 
-Most well-known global optimization methods can be formulated in accordance with this characteristical scheme, e.g.,
+Most well-known global optimization methods can be formulated in accordance with this characteristic scheme, e.g.,
 \begin{itemize}
@@
 \item a random search (Monte-Carlo) algorithm;
-\item the Piyavskii method;
-\item one-step bayesian methods proposed by Kushner and Zilinskas;
-\item information algorithms proposed by Strongin.
+\item the Piyavskii algorithm \cite{Piyavskii1972};
+\item one-step bayesian algorithms proposed by Kushner and {\v Z}ilinskas \cite{Zilinskas1989};
+\item information algorithms proposed by Strongin \cite{Strongin2000}.
 \end{itemize}
-All these methods are based on different mathematical models but are presented in the general characteristical scheme.
+All these methods are based on different mathematical models but are presented in the general characteristic scheme.
 
@@
 It is possible if the optimization method converges to the global minimum.
-
 Details of these methods are given in the sources cited. %Здесь же отметим основные моменты
+Here we highlight the main points.
 
@@
 
-The Strongin global search algorithm has been constructed in the framework of the information approach to global optimization (see \cite{Strongin2000})).
+The Strongin global search algorithm has been constructed in the framework of the information approach to global optimization (see \cite{Strongin2000}).
 This method has an adaptive scheme to evaluate a numerical estimate of the unknown Lipschitz constant.
 
-\subsection{Parallel Asynchronous Global Search Algorithm}\label{Sec_PGSA}
+\subsection{Parallel Algorithm with Asynchronous Trials}\label{Sec_PGSA}
 
@@
 
-Now, let us examine the asynchronous global search algorithm in more detail.
+Now, let us examine the parallel algorithm with asynchronous trials in more detail.
 It implements a paralleling scheme of the ``master/worker'' type.
-The master process accumulates search information, evaluates on its basis the Lipschitz constant for the objective function, determines new trial points and distributes them to the worker processes.
-Worker processes receive the points from the master, perform new trials in these points and send the results to the master.
+The master process accumulates search information, evaluates on its basis the Lipschitz constant for the target function, determines new trial points and sends them to the worker processes.
+Worker processes receive the points from the master, carry out new trials in these points and send the results to the master.
 
@@
 1) then it sends to the master process the results of the trial.
-Note that in this case we will have a set of preimages of the trial points
+Note that in this case we will have a set of inverse images of the trial points
 \[
-I_k = \left\{ x^{k+1},x^{k+2},...,x^{k+p} \right\},
+S_k = \left\{ x^{k+1},x^{k+2},...,x^{k+p} \right\},
 \]
@@
 	\item 
-Renumber the set of preimages of the trial points 
+Renumber the set of inverse images of the trial points 
 \[
@@
 \]
-containing all preimages at which the trials have either been carried out or are being carried out in ascending order (marked with subscript) , i.e.
+containing all inverse images at which the trials have either been carried out or are being carried out in ascending order (marked with subscript) , i.e.
 \[
@@
 \[
-M_1=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin I_k, x_i \notin I_k, 2\leq i\leq k+p \right\},
+M_1=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin S_k, x_i \notin S_k, 2\leq i\leq k+p \right\},
 \]
 \[
-M_2=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in I_k, 2\leq i < k+p \right\},
+M_2=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in S_k, 2\leq i < k+p \right\},
 \]
@@
 \]
-where $z_i=\varphi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$.
-The values of $z_i$ at the points $x_i \in I_k$ are undefined, as the trials at the points $x_i \in I_k$ have not yet been completed.
-If the value of $M$ equals 0, assume $M=1$.
+where $z_i=\varphi(y(x_i))$ for all $x_i \notin S_k$. 
+%The values of $z_i$ at the points $x_i \in S_k$ are undefined, as the trials at these points have not yet been completed. 
+If $M$ equals 0, then assume $M=1$.
 
 	\item
-Assign to each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, the number $R(i)$, which is called the interval characteristic and is calculated using the formula
+Assign to each interval $(x_{i-1},x_i), \; x_{i-1} \notin S_k, x_i \notin S_k$, the characteristic $R(i)$, which is calculated using the formula
 \begin{equation} \label{R}
@@
 \end{equation}
-where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.
+where $r>1$ is the method parameter and $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$.
 
 	\item
-Select the interval $[x_{t-1},x_t]$ to which the maximum characteristic corresponds, i.e.
+Find the interval $[x_{t-1},x_t]$ to which the maximum characteristic corresponds, i.e.
 \[
-R(t) = \max \left\{ R(i): \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p \right\}.
+R(t) = \max \left\{ R(i): \; x_{i-1} \notin S_k, x_i \notin S_k, \; 2\leq i\leq k+p \right\}.
 \]
 	\item
-Determine the preimage $x^{k+p+1} \in (x_{t-1},x_t)$ of the new trial point according to the formula
+Determine the inverse image $x^{k+p+1} \in (x_{t-1},x_t)$ of the new trial point according to the formula
 \[
@@
 
-Immediately after calculating the next trial point $y^{k+p+1} = y(x^{k+p+1})$, the master process adds it to the set $I_k$ and forwards it to the worker process which initiates a trial at this point.
+Immediately after calculating the next trial point $y^{k+p+1} = y(x^{k+p+1})$, the master process adds it to the set $S_k$ and forwards it to the worker process which initiates a trial at this point.
 
 The master process completes the algorithm when one of two conditions is met: $\Delta_{t}<\epsilon$, $k+p>K_{max}$.
-). The first one corresponds to stopping the algorithm by accuracy, the second one, by the number of trials. The real number $0<\epsilon<1$ and the integer $K_{max}>0$ are the algorithm parameters.
+The first one corresponds to stopping the algorithm by accuracy, the second one, by the number of trials. The real number $0<\epsilon<1$ and the integer $K_{max}>0$ are the algorithm parameters.
 
-\section{Numerical Experiments}\label{Sec_Exp}
+\section{Numerical Experiments} \label{Sec_Exp}
+
+To determine the constants of the reactions of the process, it is necessary to solve a system of differential equations describing the change in the concentration of reaction components over time \cite{cao2019}.
+
+%
 
@@
 Solving the problem with a sequential algorithm would require at least 1 million trials and thus at least 54 hours (as an estimate).
-Therefore only a parallel algorithm with a stopping criterion based on the number of trials $K_{max}=10^6$ was used for the calculations.
+Therefore, only a parallel algorithm with a stopping criterion based on the number of trials $K_{max}=10^6$ was used for the calculations.
 
 Numerical experiments were performed using the parallel asynchronous algorithm outlined in Section \ref{Sec_PGSA}, running $160$ processes: $159$ processes calculated the values of the objective function (worker processes) while $1$ master process controlled the algorithm.
-The Lobachevsky  supercomputer (University of Nizhni Novgorod)  (CentOS 7.2, SLURM, two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM on the node) was used for the experiments.
+The Lobachevsky  supercomputer (University of Nizhni Novgorod) was used for the experiments.
 The asynchronous global optimization algorithm was implemented in C++ (GCC 9.5.0 and Intel MPI were used); the computation of objective function values was implemented in Python 3.9.
+
+\begin{table}
+\caption{Dependence of the function minimum value $\varphi^*$ on the regularization parameter $\alpha$}
+\label{table_1}
+\begin{center}
+\begin{tabular}{ccc}
+\hline\noalign{\smallskip}
+ $\alpha$      & $\varphi^*$  & Time (sec.) \\
+\hline\noalign{\smallskip}
+0.0	    &	0.57392		&	838.9	\\
+0.001	&	0.54075		&	853.9	\\
+0.002	&	0.547907	&	850.0	\\
+0.003	&	0.583529	&	831.5	\\
+0.004	&	0.579725	&	821.4	\\
+0.005	&	0.615638	&	815.4	\\
+0.006	&	0.612614	&	834.7	\\
+0.007	&	0.619546	&	807.3	\\
+0.008	&	0.604817	&	830.4	\\
+0.009	&	0.63001		&	814.4	\\
+0.01	&	0.632944	&	825.2	\\
+0.02	&	0.669092	&	155.8	\\
+0.03	&	0.68152		&	809.2	\\
+0.04	&	0.736035	&	816.4	\\
+0.05	&	0.729415	&	804.7	\\
+0.07	&	0.759134	&	803.6 \\
+0.06	&	0.723404	&	810.9	\\
+0.08	&	0.808238	&	811.8	\\
+0.09	&	0.837868	&	828.6 \\
+0.1		&	0.810454	&	806.2	\\
+\noalign{\smallskip}\hline
+\end{tabular}\end{center}\end{table}
 
@@
 
-\begin{table}
-\caption{Dependence of the function minimum value $\varphi^*$ on the regularization parameter $\alpha$}
-\label{table_1}
-\begin{center}
-\begin{tabular}{ccc}
-\hline\noalign{\smallskip}
- $\alpha$      & $\varphi^*$  & Time (sec.) \\
-\hline\noalign{\smallskip}
-0.1		&	0.810454	&	806.192	\\
-0.09	&	0.837868	&	828.649 \\
-0.08	&	0.808238	&	811.852	\\
-0.07	&	0.759134	&	803.626 \\
-0.06	&	0.723404	&	810.9		\\
-0.05	&	0.729415	&	804.674	\\
-0.04	&	0.736035	&	816.387	\\
-0.03	&	0.68152		&	809.185	\\
-0.02	&	0.669092	&	155.837	\\
-0.01	&	0.632944	&	825.238	\\
-0.009	&	0.63001		&	814.48	\\
-0.008	&	0.604817	&	830.448	\\
-0.007	&	0.619546	&	807.335	\\
-0.006	&	0.612614	&	834.703	\\
-0.005	&	0.615638	&	815.45	\\
-0.004	&	0.579725	&	821.421	\\
-0.003	&	0.583529	&	831.454	\\
-0.002	&	0.547907	&	850.017	\\
-0.001	&	0.54075		&	853.898	\\
-\noalign{\smallskip}\hline
-\end{tabular}\end{center}\end{table}
+In the course of the research, the functional $f(z)$ was calculated without taking into account the addition of the regularization term (figure \ref{fig7}).
+Then a similar calculation was carried out taking into account the regularization term.
+The results for $\alpha=0.002$ is $0.527$ – with regularization, and $0.547$, respectively.
 
-\section{Conclusions and Future Work}
+\begin{figure}
+\centering
+\includegraphics[height=6.0 cm]{fig7.png}
+\caption{The value of the functional $f(z)$ without taking into account the addition of the regularization term}
+\label{fig7}
+\end{figure}
+
+\section{Conclusion}
+
+Thus, in this article an incorrect problem was solved using the Tikhonov regularization method and the reaction constants of the process of sulfuric acid alkylation of isoalkanes by alkenes were found.
+Defined fun the optimization method allowed us to find a fairly accurate description of the experimental data.
+The Tikhonov regularization method considered by the authors is planned to be used in the future to build kinetic models of other chemical processes.
+
+The use of the parallel global search algorithm showed good results.
+The solution of the problem was obtained at about 15 minutes using the Lobachevsky supercomputer, while the estimated time for solving the problem with a sequential algorithm would be more than 50 hours.
+When conducting the experiments 160 processes on the supercomputer nodes were involved in the calculations.
+However, an increase in the dimension of the global optimization problem being solved (for complex inverse problems of chemical kinetics the number of parameters can be hundreds) will lead to a decrease in the quality of the search.
+
+A possible direction for further research is the development of methods for analyzing the accumulated search information in order to identify groups of parameters that have slight effect on the objective function.
+To find the optimal values of such parameters, it is sufficient to use local optimization methods.
+Such an analysis can be carried out, for example, using machine learning methods.
+In this case, the cost of solving the problem as a whole will decrease, and it will be possible to obtain a global solution with good accuracy in an acceptable time.
 
@@
 
-\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204 and by RFBR, project No.\,19-37-60014.
+\textbf{Acknowledgments}.
+This study was supported by the Ministry of Science and Higher Education of the Russian Federation (project no.
+FSWR-2023-0034), and by the Research and Education Mathematical Center (project no.
+075-02-2022-883).
+This research was partially funded by the state task of the Institute of Petrochemistry and Catalysis of the Russian Academy of Sciences (theme No.
+FMRS-2022-0078).
 

### 2023-03-19 2023/PAVT_Ufa/PAVT2023.tex
fix one comment

@@
 	\email{leniza.enikeeva@yandex.ru} \\
+        \and
         Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia
@@
 The short time for solving the problem when using the parallel algorithm (approximately 14 minutes for 1 problem) allowed a thorough investigation of the dependence of the solution on the regularisation parameter $\alpha$.
-The table \ref{table_1} shows the minimum values of the objective function found for the corresponding value of the regularization parameter, the time for solving the problem (in seconds) and the time speedup in comparison to the estimated problem solution time when using the sequential method.
+Table \ref{table_1} shows the minimum values of the objective function found for the corresponding value of the regularization parameter and the time for solving the problem (in seconds).
+%and the time speedup in comparison to the estimated problem solution time when using the sequential method. 
 The method parameter $r=4.0$ from (\ref{R}) and the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$ were used in the runs.
@@
 Then a similar calculation was carried out taking into account the regularization term.
-The results for $\alpha=0.002$ is $0.527$ – with regularization, and $0.547$, respectively.
+The results for $\alpha=0.002$ is $0.527$ with regularization, and $0.547$, respectively.
 
@@
 \end{figure}
+
+The speedup of the parallel global optimization algorithm was estimated for the problem in which the best solution was found, i.e. the problem with regularization parameter $\alpha = 0.001$.
+Since solving the problem in serial mode would require about two days, the acceleration was evaluated in relation to a parallel launch involving a smaller number of processes.
+Table \ref{table_2} shows running time and acceleration of the parallel algorithm using 80 and 160 processes in relation to running time using 40 processes
+
+\begin{table}
+\caption{Speedup of the parallel  global optimization algorithm}
+\label{table_2}
+\begin{center}
+\begin{tabular}{ccc}
+\hline\noalign{\smallskip}
+ $p$     & Time (sec.)  & Speedup \\
+\hline\noalign{\smallskip}
+40	&	2778.7		&	---	\\
+80	&	1495.5		&	1.9	\\
+160	&	853.9	    &	3.3	\\
+\noalign{\smallskip}\hline
+\end{tabular}\end{center}\end{table}
 
@@
 
+%This research was partially funded by the Ministry of Science and Higher Education of the Russian Federation, project no.
+FSWR-2023-0034, by the Research and Education Mathematical Center, project no.
+075-02-2022-883 (development of the parallel global optimization algorithm), and by the Ministry of Science and Higher Education of the Russian Federation, project no.
+FMRS-2022-0078 (investigation of chemical kinetics problem).
+
 %
@@
 \bibliographystyle{spmpsci}
-\bibliography{bibliography}{}
+\bibliography{example-bibtex}{}
 

### 2023-03-20 2023/PAVTShort/ParallelDecisionTreeShort.tex
Короткая версия статьи по деревьям

--- 2023-02-01 2023/PAVT/ParallelDecisionTree.tex
@@
-%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% sample root file for your "contribution" to a proceedings volume
-%
-% Use this file as a template for your own input.
-%
-%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass{svproc}
-\usepackage{marvosym}
-%
-% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\usepackage{graphicx}
-
-% to typeset URLs, URIs, and DOIs
-\usepackage{url}
-\usepackage{hyperref}
-\usepackage[english]{babel} 
+\documentclass[11pt, oneside, a4paper]{article}
+%\usepackage[cp1251]{inputenc} % кодировка
+\usepackage[utf8]{inputenc} % кодировка
+\usepackage[english]{babel} % Русские и английские переносы
+\usepackage{graphicx}          % для включения графических изображений
+\usepackage{cite}              % для корректного оформления литературы
+\usepackage{enumitem}
+\usepackage{pavt-en} 
 \usepackage{amsmath}
-\def\UrlFont{\rmfamily}
-\providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}
-
-\def\orcidID#1{\unskip$^{[#1]}$}
-\def\letter{$^{\textrm{(\Letter)}}$}
 
 \begin{document}
-\mainmatter              % start of a contribution
-%
-\title{On using the decision trees to identify the local extrema in parallel global optimization algorithms}
-%
-\titlerunning{On using the decision trees to identify the local extrema }  % abbreviated title (for running head)
-%                                     also used for the TOC unless
-%                                     \toctitle is used
-%
-\author{Konstantin Barkalov\inst{1}\letter\orcidID{0000-0001-5273-2471} \and Ilya Lebedev\inst{1}\letter\orcidID{0000-0002-8736-0652} \and Dmitry Silenko\inst{1}\letter\orcidID{0000-0002-2578-9699}}
-%
-\authorrunning{Konstantin Barkalov et al.} % abbreviated author list (for running head)
-%
-%%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{K.A.~Barkalov and I.G.~Lebedev, and D.I.~Silenko}
-%
-\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
-\email{ \{konstantin.barkalov,ilya.lebedev,silenko\}@itmm.unn.ru}
-}
 
-\maketitle              % typeset the title of the contribution
+% \title - название статьи
+% \authors - список авторов
 
+\title{On using the decision trees to identify the local extrema in parallel global optimization algorithms  \footnote{This study was supported by the Russian Science Foundation, project No.\,21-11-00204 .}}
+
+\authors{Konstantin Barkalov\superscript{1},Ilya Lebedev\superscript{1}, Dmitry Silenko\superscript{1}}
+\organizations{\superscript{1}Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia}
+
+% Аннотация заключается в окружение abstract
 \begin{abstract}
@@
 This suggestion was confirmed by the results of numerical experiments demonstrating the speedup  when solving a series of test problems.
-% We would like to encourage you to list your keywords within
-% the abstract section using the \keywords{...} command.
+\end{abstract}
+
 \keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Machine learning $\cdot$ Decision tree}
-\end{abstract}
-%
+
+% \section{название} - заголовок раздела первого уровня
+% \subsection{название} - заголовок раздела второго уровня
+% \subsubsection{название} - заголовок раздела третьего уровня
+% Не используйте уровень вложенности заголовков больше трех!
+% Каждый абзац текста в статье начинается командой \par или пустой
+% строкой.
 
@@
 	\item  For each interval $(x_{i-1},x_i), 1\leq i\leq k+1,$ compute the value $R(i)$ called \textit{a  characteristic} of the interval according to the formulae
+
 	\begin{equation} 
@@
 	\end{equation}
+
 	\begin{equation} 
@@
 	\end{equation}
+
 	where \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
@@
 
-\begin{figure}[ht!]
-
-	\begin{center} 
-				\begin{minipage}[h]{0.9\linewidth} 			\includegraphics[width=1\linewidth]{figure/fig3.pdf}
-					 			\caption{Block scheme of combined GSA and Hooke-Jeeves method using the  decision tree} %% подпись к рисунку 	
-					 			\label{fig:fig3}
-				 \end{minipage}
-	\end{center}
-\end{figure}
-
 % переводить до этого раздела
@@
 As one can see from Table \ref{tab:2},  the iteration speedup is large enough for the problems of any  dimensionality.
+
+\section{Conclusion}
+
 So far, as a result of the work done, we succeeded to combine the global search algorithm with a local  optimization method.
@@
 
-\medskip
-
-\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204 .
-
 %

### 2023-03-26 2023/NUMTA/Abstract/Barkalov.tex
NUMTA abstract template

% Latex-template for one-page abstract submissions.
% NUMTA2023 - Numerical Computations: Theory and Algorithms
% 14-20 June 2023, TUI Magic Life Calabria, Pizzo Calabro (Italy)

\documentclass[oribibl]{llncs}

% Please, do not use your own macros/redefinitions.

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}
\cleardoublepage

\title{Title of Your Presentation}
\author{Name1 Surname1, Name2 Surname2}

%%% Please note that the first author is supposed to be the speaker.

\institute{University1 of Place1, Address1, Country1 \\
           University2 of Place2, Address2, Country2 \\
\email{Name1@university1.edu, Name2@university2.edu}}

\maketitle

\thispagestyle{fancy}

\textbf{Keywords.} Keyword1; keyword2; keyword3.

 \vspace*{0.5cm}

%-----------------------------------------------------------------

The goal of the Conference is to create a multidisciplinary round
table for an open discussion on numerical modeling nature by using
traditional and emerging computational paradigms. The Conference
will discuss all aspects of numerical computations and modeling from
foundations and philosophy to advanced numerical techniques. New
technological challenges and fundamental ideas from theoretical
computer science, linguistic, logic, set theory, and philosophy will
meet requirements and new fresh applications from physics,
chemistry, biology, and economy.

Researchers from both theoretical and applied sciences are very
welcome to submit their papers and to use this excellent possibility
to exchange ideas with leading scientists from different research
fields. Papers discussing new computational paradigms, relations
with foundations of mathematics, and their impact on natural
sciences are particularly solicited. A special attention will be
also dedicated to numerical optimization and different issues
related to theory and practice of the usage of infinities and
infinitesimals in numerical computations.

Papers discussing new computational paradigms and their impact on
natural sciences are particularly solicited.

\textit{Please, avoid figures and tables in the text. Max 4
bibliographic references are allowed.}

\textbf{Acknowledgements.}

This research was supported by the following grants...

 \vspace{0.5cm}

\begin{thebibliography}{4}

\bibitem{Book} Cantor G. (1955) \emph{Contributions to the founding of
the theory of transfinite numbers}. Dover Publications, New York.

\bibitem{Article} Gordon P. (2004) Numerical cognition without words:
{E}vidence from {A}mazonia. \emph{Science}, Vol.~306,
pp.~496--499.

\bibitem{Proceedings} Author I., Author I.I. (2019) An
abstract. In Proceedings of the \emph{International
Conference``XXX''} (ed. by Editor E.), Place (Country), p.~1.

\bibitem{BookChapter} Author X. (2019) A
chapter. In \emph{YYY Book} (ed. by Editor E.), Place (Country),
pp.~10--20.

\end{thebibliography}

\end{document}

### 2023-03-26 2023/NUMTA/Abstract/Barkalov.tex
full text without references

@@
 
-\title{Title of Your Presentation}
-\author{Name1 Surname1, Name2 Surname2}
+\title{Application of Machine Learning to Increase the Efficiency of the Global Search Algorithm for Solving Multicriterial Problems}
+\author{Konstantin Barkalov, Vladimir Grishagin, Evgeny Kozinov}
 
@@
 
-\institute{University1 of Place1, Address1, Country1 \\
-           University2 of Place2, Address2, Country2 \\
-\email{Name1@university1.edu, Name2@university2.edu}}
+\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
+\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, vagris@unn.ru}
 
@@
 
-\textbf{Keywords.} Keyword1; keyword2; keyword3.
+\textbf{Keywords.} Multicriterial problems; global optimization; logistic regression. % machine learning;
 
@@
 
-The goal of the Conference is to create a multidisciplinary round
-table for an open discussion on numerical modeling nature by using
-traditional and emerging computational paradigms. The Conference
-will discuss all aspects of numerical computations and modeling from
-foundations and philosophy to advanced numerical techniques. New
-technological challenges and fundamental ideas from theoretical
-computer science, linguistic, logic, set theory, and philosophy will
-meet requirements and new fresh applications from physics,
-chemistry, biology, and economy.
+Decision making models described as problems of multicriterial optimization are very complicated for investigation because of the property of criteria contradictoriness that leads to the notion of the solution as the set non-dominated parameters (Pareto set).
+The complexity of these problems increases significantly in the case where criteria are multiextremal.
+In the approaches based on convolution schemes reducing the multicriterial problem to a family of scalar (single criterion) optimization subproblems, these subproblems are multiextremal as well.
+This circumstance requires the use of global optimization methods to solve those.
 
-Researchers from both theoretical and applied sciences are very
-welcome to submit their papers and to use this excellent possibility
-to exchange ideas with leading scientists from different research
-fields. Papers discussing new computational paradigms, relations
-with foundations of mathematics, and their impact on natural
-sciences are particularly solicited. A special attention will be
-also dedicated to numerical optimization and different issues
-related to theory and practice of the usage of infinities and
-infinitesimals in numerical computations.
+In the paper a novel algorithm for multicriterial black-box optimization with multiextremal criteria are considered.
+This algorithm applies ideas of complexity reduction when the initial multicriterial problem are reduced to a set of scalar subproblems on the base of maximum convolution, then each scalar multidimensional subproblem is transformed to equivalent one-dimensional problem by means of Peano mapping which provides single-valued correspondence between N-dimensional hyperparallelepiped and the unit interval of the real axis.
+For solving univariate problems a global optimization algorithm with guaranteed convergence to global optimum is used.
+As a core novel feature, this algorithm includes in its computational scheme a machine learning procedure with combination of accumulating the information of solved subproblems that allows one to significantly accelerate the building the Pareto set.
 
-Papers discussing new computational paradigms and their impact on
-natural sciences are particularly solicited.
+For verification of workability and estimation of efficiency, the representative computational experiment on test sets of multicriterial problems with multiextremal criteria constructed on the base test class GKLS  has been conducted for different dimensionalities and number of criteria.
 
-\textit{Please, avoid figures and tables in the text. Max 4
-bibliographic references are allowed.}
+Along with the proposed algorithm, in the experiment some known method belonging to the class of genetic algorithms have been tested and compared where the hypervolume index  HV was used for assessment of algorithms' quality.
+According to the results of the experiment, the proposed algorithm  has demonstrated its essential advantage over other methods.
 
@@
 
-This research was supported by the following grants...
+This research was supported by the Russian Science Foundation, project No. 21-11-00204.
 
@@
 
-\bibitem{Book} Cantor G. (1955) \emph{Contributions to the founding of
-the theory of transfinite numbers}. Dover Publications, New York.
+\bibitem{Book} Cantor G. (1955) \emph{Contributions to the founding of the theory of transfinite numbers}. Dover Publications, New York.
 
-\bibitem{Article} Gordon P. (2004) Numerical cognition without words:
-{E}vidence from {A}mazonia. \emph{Science}, Vol.~306,
+\bibitem{Article} Gordon P. (2004) Numerical cognition without words: {E}vidence from {A}mazonia. \emph{Science}, Vol.~306,
 pp.~496--499.
-
-\bibitem{Proceedings} Author I., Author I.I. (2019) An
-abstract. In Proceedings of the \emph{International
-Conference``XXX''} (ed. by Editor E.), Place (Country), p.~1.
-
-\bibitem{BookChapter} Author X. (2019) A
-chapter. In \emph{YYY Book} (ed. by Editor E.), Place (Country),
-pp.~10--20.
 

### 2023-04-06 2023/NUMTA/Abstract/Barkalov.tex

@@
 
-\bibitem{Book} Cantor G. (1955) \emph{Contributions to the founding of the theory of transfinite numbers}. Dover Publications, New York.
+\bibitem{Sergeyev2013}
+Sergeyev Y.D., Strongin R.G., Lera D. (2013) \emph{Introduction to global optimization exploiting space-filling curves}. Springer Briefs in Optimization, Springer, New York.
 
-\bibitem{Article} Gordon P. (2004) Numerical cognition without words: {E}vidence from {A}mazonia. \emph{Science}, Vol.~306,
-pp.~496--499.
+\bibitem{Gergel2018}
+Gergel V.P., Kozinov,E.A. (2018) Efficient multicriterial optimization based on intensive reuse of search information. \emph{J. Glob. Optim.},  Vol. 71, pp.~73--90.
 

### 2023-04-06 2023/PAVT_poster/poster.tex

--- 2022-03-24 2022/PAVTOneAPI/PAVT_Poster/poster.tex
@@
             style=gost-numeric,
-            sorting=nty,
+            sorting=none,
             autolang=other]{biblatex}
@@
 
-\title{Решение многомерных задач глобальной оптимизации с использованием Intel oneAPI }
-\author{К.А. Баркалов \and И.Г. Лебедев \and Я.В. Кольтюшкина}
+\title{Параллельный алгоритм липшицевой глобальной оптимизации с двойной оценкой константы Липшица }
+\author{К.А. Баркалов \and  Д.А. Черных}
 \institute{Нижегородский государственный университет им. Н.И. Лобачевского}
@@
             \setlength{\parindent}{1.25cm} %
-             Задача многомерной многоэкстремальной оптимизации может быть определена как поиск наименьшего значения действительной функции \(\phi(y)\)  в гиперинтервале \(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
+            Рассмотрим задачу глобальной оптимизации вида\\
+             \(\phi(y^*) = min\{\phi(y), y\in D\}\), где \(D=\{y\in R^N:a_i\leqslant y_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\).
 
-             Будем предполагать, что целевая функция задана по принципу <<черного ящика>> и удовлетворяет условию Липшица с априори неизвестной константой \(L\).
-
+             Будем предполагать, что целевая функция задана как <<черный ящик>> и удовлетворяет условию Липшица с априори неизвестной константой \(L\) \cite{Kvasov2013}. 
           \end{block}
@@
                   %\centering
-                  \center{\includegraphics[width=1.01\textwidth]{images/pean_2 (1)_8.jpg}}
+                  \center{\includegraphics[width=1.000\textwidth]{images/pean_2 (1)_8.jpg}}
                   %\includegraphics[scale=1.468]{images/pean_2.jpg}
@@
 \end{block}
-\begin{block}{Метод глобальной оптимизации}
+\begin{block}{Метод с  двойной оценкой константы Липшица}
 \leftskip=0.5cm \rightskip=0.5cm 
 \setlength{\parindent}{1.25cm} %
-В процессе работы алгоритма строится последовательность точек \(x^k\), в которых проводятся \textit{испытания}, т.е. вычисляются значения целевой функции \(\phi^k=\phi(y(x^k))\).
+Алгоритм строит последовательность точек \(x^k\), в которых проводятся \textit{испытания}, т.е. вычисляются значения целевой функции \(\phi^k=\phi(y(x^k))\).
+
+В исследуемом методе для достижения более точного значения константы \(L\) используются заниженная и завышенная оценки константы Липшица\cite{Strongin2020}.
+Большая оценка гарантирует глобальную сходимость алгоритма, меньшая снижает число поисковых испытаний, требующихся для поиска решения задачи.
 
@@
 Поэтому параллельное проведение нескольких испытаний повышает скорость работы алгоритма.
-
 \leftskip=0.001cm 
@@
                  % \centering
-                 \center{\includegraphics[width=1.045\textwidth]{images/PGS2.eps}}
+                 \center{\includegraphics[width=1.045\textwidth]
+                 {images/PGS2.eps}}
                 % \includegraphics[scale=1.095]{images/PlotGS_7.png}
@@
                 \item Упорядочить точки предшествующих испытаний в порядке возрастания их координат: \(x_{0}<...<x_{i}<...<x_{k}\);
-                \item Для каждого интервала \((x_{i-1}, x_{i}),1\leqslant i\leqslant k,\) вычислить его \textit{характеристику} \(R(i)\);
+                \item Для каждого интервала \((x_{i-1}, x_{i}),1\leqslant i\leqslant k,\) вычислить две \textit{характеристики} \(R_{glob}(i), R_{loc}(i)\), отвечающие за завышенную и заниженную оценки константы Липшица;
+                \item Вычислить характеристику \(R(i) = \max⁡\{\rho*R_{loc}(i),R_{glob}(i)\}\), где \(\rho\) -- весовой коэффициент; 
+                % \(\rho=((1-1/r_{glob} )/(1-1/r_{loc} ))^2\)
                 \item Упорядочить характеристики \(R(i)\) в порядке убывания \(R(t_{1})\geqslant R(t_{2})\geqslant ...
@@
 
-\begin{block}{Использование Intel oneAPI}
-
-\leftskip=0.5cm \rightskip=0.5cm 
-\setlength{\parindent}{1.25cm} %
-Мы реализовали рассмотренный параллельный алгоритм с использованием набора инструментов Intel oneAPI.
-Достоинствами Intel oneAPI являются простота, открытость и возможность легкой интеграции в существующий код.
-Intel oneAPI позволяет написать код один раз и в дальнейшем запускать его на различных устройствах.
-
-При реализации мы использовали возможности распараллеливания, которые обеспечиваются языком Data Parallel C++ и набором библиотек Intel oneAPI, облегчающих межархитектурную разработку.
-
-\end{block}
 %----------------------------------------------------------------------------------------
@@
 
-\begin{block}{Результаты}
+\begin{block}{Результаты экспериментов с использованием OpenMP}
 
@@
 \setlength{\parindent}{1.25cm} %
+При реализации параллельного алгоритма использовалась технология \textit{OpenMP} для программирования многопоточных приложений на многопроцессорных системах с общей памятью.
+
+Достоинствами \textit{OpenMP} являются простота, открытость, возможность легкой интеграции в существующий код, а так же возможность инкрементального программирования.
+
 На данный момент получены результаты вычислительных экспериментов на тестовых задачах, формируемых специальным генератором \cite{GKLS}.
-В таблицах \ref{table:GKLS_RES_1}, \ref{table:GKLS_RES_2} приведены среднее число итераций параллельного алгоритма глобального поиска и ускорение относительно последовательной версии.
-В каждом эксперименте решалось 100 задач размерности \(N = 4\) и \(N = 5\).
-Эксперименты проведены на компьютере с процессором Intel Core I5-7300HQ 2.5 GHz с видеокартой HD Graphics 630, 16 GB RAM, использовался язык \textit{Data Parallel C++}.
+В таблице \ref{table:GKLS_RES_1} приведены среднее число итераций параллельной реализации метода с двойной оценкой константы Липшица и ускорение относительно последовательной версии.
+В эксперименте решалось 100 задач размерности \(N = 2\).
+Результаты получены на узле суперкомпьютера <<Лобачевский>> (процессор AMD EPYC 7742 64-Core, 40 GB RAM, видеокарта NVIDIA GA100), использовался язык \textit{C++} с технологией \textit{OpenMP} .
 
@@
     \centering
-    \caption{Результаты решения серии задач размерности $N=4$}
+    \caption{Результаты решения серии задач размерности $N=2$}
      \renewcommand{\arraystretch}{1.4}
@@
     P  & Итерации  & Время работы (сек.) & Ускорение \\ \hline
-	128 & 106.5 & 1.2  & 4.7    \\ \hline
-	256 & 41.3 & 0.66   & 8.5          \\ \hline
+        16 & 15.2 & 6.52 & 10.87  \\ \hline
+	32 & 8.12 & 3.75 & 18.9  \\ \hline
+	64 & 5.04 & 3.00 & 23.62 \\ \hline
 	\end{tabular}
@@
     \label{table:GKLS_RES_1}
-\end{table}
-
-\begin{table}[!hbp]
-    \centering
-    \caption{Результаты решения серии задач размерности $N=5$}
-    \renewcommand{\arraystretch}{1.4}
-    \renewcommand{\tabcolsep}{1cm}
-    \begin{tabular}{|c|c|c|c|}
-    \hline
-    P  & Итерации  & Время работы (сек.) & Ускорение \\ \hline
-	128 &        235.6 & 2.6  & 4.4      \\ \hline
-	256 &       90.5  & 1.2 & 9.2      \\ \hline
-	\end{tabular}
-
-    \label{table:GKLS_RES_2}
 \end{table}
@@
             %\bibliography{bibliography}
-
             \printbibliography

### 2023-06-09 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
Corrected vestion of NUMTA2023 paper

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{cite}
\usepackage{hyperref}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
 \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Application of Machine Learning to Increase the Efficiency of the Global Search Algorithm for Solving Multicriterial Problems\thanks{This research was supported by the Russian Science Foundation, project No. 21-11-00204.}}
%
\titlerunning{Global Search Algorithm for Solving Multicriterial Problems}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
Vladimir Grishagin\orcidID{0000-0002-2884-3670} \and
Evgeny Kozinov\orcidID{0000-0001-6776-0096}}
%
\authorrunning{K. Barkalov, V. Grishagin and E. Kozinov}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, \email{vagris@unn.ru}}

%
\maketitle              % typeset the header of the contribution
%
	\begin{abstract}
Decision making models described as problems of multicriterial optimization are very complicated for investigation because the property of criteria contradictoriness leads to the notion of the solution as the set of non-dominated parameters (Pareto set). The complexity of these problems increases significantly in the case where criteria are multiextremal.
In the paper a novel algorithm for multicriterial black-box optimization with multiextremal criteria is considered. This algorithm applies ideas of complexity reduction when the initial multicriterial problem is reduced to a set of univariate scalar subproblems on the base of Peano mapping and maximum convolution. For solving univariate problems an information-statistical global optimization algorithm with guaranteed convergence to global optimum is used. As a core of novelty, this algorithm includes in its computational scheme a machine learning procedure with combination of accumulating the information of solved subproblems that allows one to significantly accelerate building the Pareto set.
Effectiveness of the proposed approach is estimated in the representative computational experiment on test sets of multicriterial problems with multiextremal criteria for different dimensions and number of criteria in comparison with several nature-inspired multiobjective optimization algorithms.

\keywords{Multicriterial problems \and Global optimization \and Logistic regression}
\end{abstract}
%
%
%
\section{Introduction}
Artificial intelligence models formulated as decision-making problems are extremely diverse and attract significant attention of researchers due to the richness of theoretical background and practical importance. The high complexity of decision-making models, along with the use of existing effective approaches to their analysis, requires the development of new ideas that overcome these difficulties. A promising way to improve decision-making algorithms is to combine in their development the advantages of different approaches. For example, machine learning procedures effectively use built-in optimization methods to enhance the quality of learning.

In present paper we, on the contrary, propose an approach in which a machine learning technique embedded into optimization algorithm allows one to increase its efficiency when this algorithm is applied to solving complicated multiobjective problems.

As the decision-making model we consider the multicriterial multidimensional optimization problems with multiextremal ``black-box'' criteria herein after called as MMCO problems. Combination of multicriteriality and multiextremality generates the dramatic complexity of these problems, because, first, criteria are contradictory, as a rule, and this circumstance leads to a complicated concept of solution (Pareto set) in such problems and, second, in the multiextremal case the growth of computational efforts is exponential when increasing the dimensionality (number of independent variables).

For analyzing multiobjective problems under consideration the following general approaches can be used. One of them is based on placement of a grid (deterministic or random) in the feasible domain, pairwise comparison of grid nodes in accordance with the domination principle and selection of non-dominated points.

The second approach applies meta-heuristic methods (genetic algorithms, simulation annealing, particle swarm optimization, artificial bee or ant colony, firefly algorithm, etc.) inspired by modeling the behavior of biological agents \cite{Nebro2009,RC05,Mostaghim2007,Durillo2010} or physical processes \cite{DPA02,ZLT01}. Methods of this class are simple enough for implementation, but because of heuristic nature they do not guarantee convergence to global optimum in multiextremal case.

The third approach is based on the reduction of multicriterial problem to a family of single-criterion (scalar) optimization subproblems by means of convolution method and solving these subproblems by efficient mathematical programming algorithms \cite{Collette2004,Marler2009,Marler2004,Pardalos2017,Wierzbicki,Ehrgott2005,GergelKozinov2020}. Under corresponding conditions, the solution of the scalar subproblem is a Pareto-optimal point. When changing coefficients of the convolution it is possible to find an approximation of the Pareto set. If the criteria in the initial multiobjective statement are multiextremal, the reduced convolution function is multiextremal as well, and for its optimization it is necessary to apply search algorithms that provide the finding of the global optimum.

In the class of deterministic methods for multiextremal optimization with guaranteed convergence to global optimum \cite{Gablonsky2001,Sergeyev2010,Sergeyev2006,Evtushenko2009,Evtushenko2014,Barkalov2015,Strongin2003,Sergeyev2013}, the algorithms developed in the framework of the information-statistical approach (ISA) \cite{Gergel2017_1,Gergel2018,Barkalov2014,Barkalov2018} are among the most effective. The present paper describes a new algorithm belonging to the ISA class and aimed at solving MMCO problems that accelerates the Pareto set building by the application of machine learning ideas.

The proposed method sequentially implements the concept of complexity reduction. Following this way, the initial MMCO problem is reduced to a family of single-criterion optimization subproblems by means of parameterized convolution. Then each multidimensional subproblem is transformed to an equivalent univariate (one-dimensional) optimization problem formed on the base of Peano-type mapping \cite{Evtushenko2014,Gergel2017_1,Gergel2018,Marler2004,Pardalos2017}. Further, a special global optimization algorithm designed on the ISA platform is applied for solving the univariate problems. Its computational scheme includes two accelerators that decreases significantly number of algorithm's iterations. The first one when solving current one-dimensional problem takes into account results of iterations of all the previous problems solved already. The second mechanism influences the choice of the new iteration point estimating possible closeness of its pre-image in the criteria space to the Pareto set. This estimation is carried out by the method of separation planes, which is widely used in machine learning.

For efficiency confirmation of the proposed approach combining the global optimization algorithm with machine learning procedures, the representative computational experiment has been conducted on the test class of multiextremal multicriterial problems of different dimensions and number of criteria. For expanding the confirmation area, the test problems were solved by several known meta-heuristic methods of multicriterial optimization. The Hyper-Volume (HV) index reflecting the quality of the Pareto set approximation was used as a comparison criterion. The results of the experiment have demonstrated significant advantages of the new algorithm over other methods having involved in comparison.

The rest of the paper is organized as follows. Section \ref{sec:2} contains statement of the problem to be considered. Section \ref{sec:3} is devoted to the description of the proposed algorithmic schemes.  Results of computational experiment are presented in Section \ref{sec:4}. Section \ref{sec:5} concludes the article.

\section{Problem statement}\label{sec:2}

The decision making model to be considered hereinafter as the MMCO problem can be described in the following statement.

Let real-valued functions $\varphi_s(y)$, $1 \leq s \leq k$, depend on vector of independent variables 
$y=(y_1, y_2, \dots, y_n)$ and are defined over a box-domain (hyperparallelepiped)
\begin{equation}
\label{eq:01}
    H=\{y \in R^n : a_i \leq y_i \leq b_i, 1 \leq i \leq n\}
\end{equation}                                      in $n$-dimensional Euclidean space $R^n$.

In decision making terminology the region (\ref{eq:01}) is called \textit{feasible domain}, the argument $y$ belonging to the domain (\ref{eq:01}) is \textit{feasible decision}.  Functions  $\varphi_s(y)$, $1 \leq s \leq k$, are called \textit{partial criteria} of the problem. These functions can be interpreted as goals of decision making and, for certainty, we will assume that the less the function value, the better the result of achieving this goal.

Partial criteria form a vector-function called \textit{vector criterion}
\begin{equation}
\label{eq:02}
    \Phi(y) = (\varphi_1(y), \varphi_2(y), \dots, \varphi_k(y))
\end{equation}
and problem of multicriterial optimization is formulated as the problem of minimizing the vector function (\ref{eq:02}) over the domain (\ref{eq:01}):

\begin{equation}
\label{eq:03}
    \Phi(y) \to \min, y \in H.
\end{equation}

Unfortunately, finding a decision   that provides the minimal values for all partial criteria simultaneously is possible only in exceptional cases because, as a rule, partial criteria are contradictory, when decreasing one criterion leads to increasing the other.

For defining the solution in this case, the known concept of domination is introduced to exclude obviously unsatisfactory decisions \cite{Pardalos2017,Marler2009}.
\begin{definition}
Let $y, z \in H$. Vector $y$ is said to dominate vector $z (y \succ z)$ if  
\begin{equation}
    \label{eq:04}
    \varphi_s(y) \leq \varphi_s(z), 1 \leq s \leq k
\end{equation}
and there exists a number $t$, $1 \leq t \leq k$, such that $\varphi_t(y) < \varphi_t(z)$.
\end{definition}

After that, the set of non-dominated (efficient, Pareto optimal) points is considered as the solution to the multicriterial problem. Let's reproduce the strict relevant definitions.
\begin{definition}
A decision vector $y^* \in H$ is a Pareto optimal point if in the domain $H$ there is no vector $z, z\neq y^*$ dominating $y^*$.
\end{definition}
\begin{definition}
The set of all Pareto optimal points (Pareto set $P$) is the Pareto optimal solution to the MMCO problem (\ref{eq:03}).
\end{definition}

There exist different approaches to searching for Pareto points, briefly discussed in Introduction. In our research we have applied the convolution approach when the multicriterial problem is reduced to a family of scalar (single-criterion) problems such that the solutions to these problems are Pareto optimal points.

The parameterized function
\begin{equation}
    \label{eq:05}
    V(\mu, y) = \max_{1 \leq s \leq k} \mu_s \varphi_s(y)
\end{equation}
is taken as the convolution (so called \textit{maximum convolution}) where parameters $\mu_s, 1 \leq s \leq k$, (\textit{weight coefficients}) are such that 
\begin{equation}
    \label{eq:06}
    \mu_s \geq 0, 1 \leq s \leq k, \sum_{s=1}^k \mu_s=1.
\end{equation}
Vector
\begin{equation}
    \label{eq:06a}
    \mu = (\mu_1, \dots, \mu_k) 
\end{equation}
under conditions (\ref{eq:06}) belongs to a $(k-1)$-dimensional convex polyhedron in the Euclidean space  $R^k$.

If in the domain $H$ all the criteria 
\begin{equation}
    \label{eq:07}
    \varphi_s(y) > 0, 1 \leq s \leq k,
\end{equation}
then global minimizer of the problem
\begin{equation}
    \label{eq:08}
    W(\mu, y) = V(\mu, y) + \alpha \sum_{s=1}^k \mu_s \varphi_s(y) \to \min, \; y \in H,
\end{equation}
is a Pareto point of the problem (\ref{eq:03}) ($\alpha$  is a sufficiently small parameter) \cite{Wierzbicki,Marler2004}.

If one or more partial criteria are not positive, this is not too restrictive, because it is possible to convert this problem to the statement satisfying conditions (\ref{eq:07}) with preserving the Pareto solution.

Theoretically, for constructing the complete Pareto set of a problem (\ref{eq:02}) it is sufficient to solve problems (\ref{eq:08}) with all possible weight coefficients (vectors $\mu$ from polyhedron (\ref{eq:06})). However, when numerical solving we can deal with a finite number of problems (\ref{eq:08}) and, therefore, obtain only an approximation of the Pareto set. In these circumstances, important aspects are choice of weight coefficient and evaluation of approximation quality.

In our research we took weight coefficients as nodes of a uniform grid in the polyhedron (\ref{eq:06}). For assessment of the quality of Pareto set approximation obtained after solving problems (\ref{eq:08}) with mentioned weights, the HyperVolume index (HV-index) \cite{Evtushenko2014,Gergel2017_1,Gergel2018} was applied as traditional criterion of approximation accuracy.

\section{Global Search Algorithm for Solving MMCO Problems}\label{sec:3}

\subsection{Preliminary Information}

In the framework of our consideration, the partial criteria $\varphi_s(y)$, $1 \leq s \leq k$, are supposed to be ``black-box'' functions and to satisfy in the feasible domain (\ref{eq:01}) the Lipschitz condition
\begin{equation}
    \label{eq:09}
    |\varphi_s(y') - \varphi_s(y'')| \leq L_s \|y'-y''\|, \; y', y'' \in H, \; 1 \leq s \leq k,
\end{equation}
where constant values $L_s$, $1 \leq s \leq k$, (Lipschitz constants) are usually unknown. Notation $\| \cdot \|$ represents the Euclidean norm in $R^n$.

In general case, the functions satisfying the Lipschitz condition are multiextremal, therefore, the convolution (\ref{eq:05}) will be multiextremal and Lipschitzian as well. It means that for solving the problem (\ref{eq:08}) it is required to apply methods of non-linear programming which provides finding the global optimum. There are many global optimization algorithms designed under different assumptions about properties of objective function; relevant references can be found, for instance, in monographs \cite{Pardalos2017,Marler2009,Sergeyev2013}. Some of them are theoretically substantiated and guarantee the convergence to global solution; the others are empirical, for example, nature-inspired methods that do not ensure finding the global optimum.

The information-statistical algorithms \cite{Gergel2017_1,Gergel2018,Barkalov2014,Barkalov2015,Barkalov2018,Strongin2003}  as deterministic methods with guaranteed convergence are among the most effective for analyzing multiextremal optimization problems. For solving multidimensional problems, ISA algorithms apply two approaches based on ideas of dimensionality reduction. In the framework of one approach the multidimensional problem is reduced to a family of univariate problems connected recursively (scheme of nested optimization) \cite{Grishagin2016,Zilinskas2015,Gergel2019_2}. The other approach \cite{Gergel2018,Barkalov2014,Sergeyev2013} exploits the known fact that $n$-dimensional box (1) and the closed interval $[0,1]$ of the real axis are equipotent sets and there exist unambiguous and continuous mappings of this interval onto the hyperparallelepiped (1) called \textit{Peano-type} curves or \textit{evolvents}.

Existence of Peano mappings can be applied to solving the optimization problems in the following way. Let there be a $n$-dimensional continuous function $g(y)$ defined in the domain (\ref{eq:01}) and $y(x)$ is a Peano-type curve. Then the continuity of $g(y)$ and $y(x)$ allows one to assert that
\begin{equation}
    \min_{y\in H} g(y) = \min_{x\in[0,1]} g(y(x))
\end{equation}
i.e., instead of minimization of the multidimensional function $g(y)$ in the domain (\ref{eq:01}) one can solve the minimization problem for the one-dimensional function $g(y(x))$ over interval $[0,1]$.

However, because of fundamental characteristics of the mapping $y(x)$ the reduced function $g(y(x))$ does not retain Lipschitz property in the case of Lipschitzian partial criteria from (\ref{eq:02}). In this situation the function $g(y(x))$ satisfies the H{\" o}lder condition 
\begin{equation}
    \label{eq:10}
    |g(y(x')) - g(y(x''))| \leq G \|x'-x''\|, x', x'' \in [0,1]
\end{equation}
where $G>0$ is H{\" o}lder constant. If the function $g(y)$ satisfies the Lipschitz property with a constant $L$ then the function $g(y(x))$ will satisfy the inequality (\ref{eq:10}) with $G=2L\sqrt{n+3}$ \cite{Strongin2003,Sergeyev2013}.

Before starting the general description of the proposed algorithm, let's introduce some designations.

Let the set $M$ contain the finite number $m$ of different weight vectors (\ref{eq:06a}):
\begin{equation}
    \label{eq:11}
    M=(\mu^1, \dots, \mu^m)
\end{equation}
where
\begin{equation}
    \label{eq:12}
    \mu^i=(\mu^i_k, \dots, \mu^i_k), 1 \leq i \leq m.
\end{equation}

The set $M$ generates a family   of scalar problems 
\begin{equation}
    \label{eq:13}
    W(\mu^i, y)= V(\mu^i, y) + \alpha \sum^k_{s=1}\mu^i_s \varphi_s (y) \to \min, y \in H, 1 \leq i \leq m.
\end{equation}

Let's denote as $f^i(x)$ the univariate function obtained in accordance with Peano mapping $y(x)$ from the function $W(\mu^i, y)$, i.e.,
\begin{equation}
    \label{eq:14}
    f^i(x) = W(\mu^i, y(x)), 1 \leq i \leq m.
\end{equation}

Minimization of a separate function $f^i(x)$ over interval $[0,1]$ allows one to find partial solution(s) of the initial MMCO problem (\ref{eq:03}) and solving all the problems
\begin{equation}
    \label{eq:15}
    f^i(x) \to min, x \in [0,1], 1 \leq i \leq m,
\end{equation}
enables to construct an approximation of the Pareto set.

The following computational scheme is proposed for solving the problems of the family (\ref{eq:15}). In this scheme, all the problems (\ref{eq:15}) are solved sequentially and during minimization of current function $f^i(x), i > 1$, the results of preceding problems $f^j(x) \to \min, 1 \leq j \leq i - 1$ completed already are taken into account.

Before proceeding to the detailed description of the scheme, we will make some preliminary comments.

Hereinafter the term ``trial'' will denote the evaluation of the objective function $f^i(x)$ at a point $x^j \in [0,1]$ and obtaining the vector (result of trial)
\begin{equation}
    \label{eq:15a}
    \Omega^i(x^j)=(z^{i,j}, \omega^j_1, \dots, \omega^j_k),
\end{equation}
where $z^{i, j}=f^i(x^j)$, and the quantities
\begin{equation}
    \label{eq:16}
    \omega^i_s=\varphi_s(y(x^j)), 1 \leq s \leq k,
\end{equation}
are values of corresponding partial criteria at the point $x^j$. If a vector $M$ of weight coefficients (\ref{eq:11}) corresponds to the function (\ref{eq:14}) then 
\begin{equation}
    \label{eq:17}
    z^{i, j} = \max_{1\leq s \leq k} \mu_s \omega^j_s + \alpha \sum^k_{s=1} \omega^j_s.
\end{equation}

It should be noted that if the value (\ref{eq:17}) has been already calculated and the vector $\Omega^i(x^j)$ has been stored, then evaluation of another function with its weight coefficient set $M'=(\mu'^1, \dots, \mu'^k)$ at the point $x^j$ does not require the recalculation the partial criteria (\ref{eq:02}) because the value $z^{l, j}$ can be easily computed according to (\ref{eq:17}).

For this goal, it is useful to organize a special matrix of a posteriori information which will store results of all conducted trials, for example, in the form
\begin{equation}
    \label{eq:18}
A^i_p =
\begin{pmatrix}
x^1 & x^2 & \dots & x^p \\
z^{i,1} & z^{i,2} & \dots & z^{i,p} \\
\omega_1^{i,1} & \omega_1^{i,2} & \dots & \omega_1^{i,p} \\
\dots & \dots & \dots & \dots \\
\omega_k^{i,1} & \omega_k^{i,2} & \dots & \omega_k^{i,p}
\end{pmatrix}
\end{equation}
after implementing $p$ trials at points $x^1, x^2, \dots, x^p$ when minimizing a function $f'(x)$ from (\ref{eq:15}).

\subsection{General Computational Scheme}

Before starting the described algorithm to build a Pareto set approximation, the set (\ref{eq:11}) of weight coefficient vectors is supposed to be given. In the framework of the statement (\ref{eq:01})-(\ref{eq:03}) this set generates the family of $m$ multidimensional scalar problems (\ref{eq:13}) and corresponding to those the family of $m$ univariate problems (\ref{eq:13}).

The general algorithm consists of $m$ stages. At the $i$-th stage, solving the problem (\ref{eq:15}) for the function $f^i(x)$  is carried out.

At the beginning of each stage initial trials are implemented in $q$ given points $x^1, x^2, \dots, x^q \in [0,1]$.

At the first stage when $f^i(x)=f^1(x)$, $q=2$ and $x^1=0, x^2=1$. These 2 trials generate results $\Omega^1(x^1), \Omega^1(x^2)$ from (\ref{eq:15a}) and initial information matrix $A^1_2$ from (\ref{eq:18}).

In all the stages after the first one $(i>1)$, the trial coordinates obtained in all previous stages are taken as points of initial trials at current stage. As the matrix $A^{i-1}_q$ contains values of all partial criteria at these points it is easy to recalculate values $z^{i,1}, z^{i,2}, \dots, z^{i,p}$ of function $f^i(x)$ according to (\ref{eq:17}) and to build the updated matrix $A^i_q$.

Further computational procedure can be described in the following manner.

Let after completing $p \geq q$ trials at points $x^1, x^2, \dots, x^p$ the matrix $A^i_p$ has been obtained. For computing a new trial point $x^{p+1}$ it is necessary to implement operations in accordance with the following rules.

\textbf{Rule 1.} Points $x^j, 1 \leq j\leq p,$ are ordered in increasing order and renumbered by subscripts, i.e.,
\begin{equation}
    \label{eq:19}
    x_0 = 0 < x_1 < \dots < x_{p-2} < x_{p-1} = 1.
\end{equation}

The values $u_j = f^i(x_j), 1\leq j \leq p$ are juxtaposed to the points $x_j, 1\leq j \leq p$.

\textbf{Rule 2.} Current estimation $\gamma$ of the H{\" o}lder constant  
\begin{equation}
    \label{eq:20}
\gamma=
\begin{cases}
  \begin{matrix}
     \gamma \Gamma & \Gamma >0 \\
     1 & \Gamma = 0 
  \end{matrix}
\end{cases}
\end{equation}
is computed, where
\begin{equation}
    \label{eq:21}
\Gamma=
\max_{1 \leq j \leq p-1} {\frac{|u_j - u_{j-1}|}{\rho_j} },
\end{equation}
\begin{equation}
    \label{eq:22}
\rho_j = \sqrt[n]{x_j-x_{j-1}},
\end{equation}
and $r>1$ is a parameter of the method.

\textbf{Rule 3.} In the set of trial coordinates $x_j, 1 \leq j \leq p$, non-dominated points are selected according the rule of Definition 2 and all such points are considered as points belonging to current approximation of the Pareto set. All columns of the information matrix $A^i_p$ are marked up with labels 0 or 1 where 1 corresponds to non-dominated points and 0 to dominated ones.

\textbf{Rule 4.} In the criteria space on the base of the marked up matrix $A^i_p$ a separating hyperplane $L_p$ is constructed applying logistic regression \cite{Yu2011} as efficient tool of machine learning.

\textbf{Rule 5.} For each point $x_j, 1 \leq j \leq p$, the distance $d_j$ of the vector $(\varphi_1(x_j), \dots, \varphi_k(x_j))$ to the separating plane $L_p$ is determined. This distance is considered to be negative if vector is placed above the plane and positive otherwise (see Fig. \ref{fig:1}).

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{fig1}
\caption{An example of estimating distances to the separating plane $L_p$.} \label{fig:1}
\end{figure}

\textbf{Rule 6.} Distances $d_j, 1 \leq j \leq p$, are scaled:     
\begin{equation}
    \label{eq:23}
d'_j=
\begin{cases}
  \begin{matrix}
     d_j / d_{max}, & d_j > 0 \\
     -d_j / d_{min}, & d_j < 0 
  \end{matrix}
\end{cases}
\end{equation}
where $d_{max} = \max_{1 \leq j \leq p} {d_j}$, $d_{min} = \min_{1 \leq j \leq p} {d_j}$.

\textbf{Rule 7.} Neighboring points from (\ref{eq:19}) provide a partitioning of the interval $[0, 1]$ into $p-1$ subintervals $(x_{j-1}, x_j), 1 \leq j \leq p-1$.

A real value $B(j)$ is assigned to each subinterval $(x_{j-1}, x_j), 1 \leq j \leq p-1$. $B(j)$ is called \textit{the characteristic}  of this subinterval and is calculated as
\begin{equation}
    \label{eq:24}
    B(j) = \rho_j + \frac{(u_j-u_{j-1})^2}{\gamma^2 \rho_j} - \frac{2 (u_j+u_{j-1})}{\gamma} + \beta (d'_j + d'_{j-1})
\end{equation}
where $\beta > 0$ is a parameter influencing the contribution of the separating plane.

\textbf{Rule 8.} Among all subintervals, the subinterval $(x_{j-1}, x_j), 1 \leq j \leq p-1$, with the highest characteristic is chosen:
\begin{equation}
    \label{eq:25}
    B(l) = \max_{1 \leq j \leq p-1} {B(j)}.
\end{equation}

\textbf{Rule 9.} The new $(p+1)$-th trial is executed at the point  
\begin{equation}
    \label{eq:26}
    x^{p+1} = \frac{x_l + x_{l-1}}{2} - sign(u_l - u_{l-1}) \frac{1}{2r} \left(\frac{|u_l - u_{l-1}|}{\Gamma} \right)^n,
\end{equation}
values $\omega^{p+1}_s, 1 \leq s \leq k$, and $z^{i,p+1}=f^i(x^{p+1})$  are calculated and along with the point (\ref{eq:26}) they form the new matrix $A^i_{p+1}$ adding new column to the matrix $A^i_p$.

\textbf{Rule 10.} If the inequality
\begin{equation}
    \label{eq:27}
    \rho_l=\sqrt[n]{x_l-x_{l-1}} < \varepsilon
\end{equation}
is true where $\varepsilon > 0$ is predefined accuracy, then the $i$-th  stage (minimization of function $f^i(x)$) is over.

In the case $i < m$, the method proceeds to the next stage $(i = i+1)$.

If $i=m$, then the algorithm completes its running and the set of non-dominated trial points in the criteria space is taken as a Pareto set approximation.

\section{Experimental efficiency testing}\label{sec:4}

Computations were conducted on the supercomputer ``Lobachevsky'' of the Nizhni Novgorod State University (operating system CentOS 7, control system SLURM). One node of the supercomputer has 2 processors Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM. One CPU contains 8 cores, i.e., 16 cores are available on one node. To obtain the executable program code the compilers Intel C++ 17.0.0 and python 3.9 were used. Computations were performed with the support of the package scikit-learn 0.24.2 and the applied optimization system Globalizer \cite{globalizerSystem}.

The experiment was organized in the following way. For efficiency evaluation of the proposed algorithm hereinafter called AGS-SP (Algorithm of Global Search with Separating Planes), it was compared with its modification AGS-R (Algorithm of Global Search – Reduced) where computational scheme is similar, but Rules 3- 6 are removed and in the characteristic (\ref{eq:24}) parameter $\beta = 0$, AGS-R ignores the information about the Pareto set approximation. Moreover, to expand the basis of the comparison, some known meta-heuristic evolutionary MMCO methods were involved into the experiment, namely, algorithms SMPSO \cite{NDG09}, OMOPSO \cite{RC05}, NSGA-II \cite{DPA02}, SPEA2 \cite{ZLT01}, IBEA \cite{ZK04} from the library jMetal \cite{DNA10}.

As a criterion of effectiveness, the HyperVolume (HV) index \cite{Evtushenko2014,Gergel2018} were adopted. It is traditionally used to estimate the efficiency of the quality for Pareto set approximations built by numerical methods. HV-index characterizes the completeness of the approximation (a larger value corresponds to a more complete coverage of the Pareto region). To compare correctly the relative effectiveness of different methods in a unified scale the ``ideal'' value of HV index was calculated on a dense grid in the feasible domain for each test MMCO problem

Three series of experiments were executed. The first one dealt with a simple problem and was aimed at demonstration of positive influence of criteria space analysis by means of separating planes and at comparing with evolutionary techniques.

This problem was bi-criteria one \cite{CHIANDUSSI2012912}:
\begin{equation}
    \label{eq:28}
    \begin{matrix}
f_1 (y)=y_1 \\
f_2 (y)=(1+10y_2 )[1- \left(\frac{y_1}{1+10y_2} \right)^2- \frac{y_1}{1+10y_2} \sin(8 \pi y_1 ) ] \\
0 \leq y_1,y_2 \leq 1
    \end{matrix}
\end{equation}

AGS-SP and AGS-R solved 20 scalar problems (\ref{eq:08}) for different weight coefficients $\mu_1$ uniformly distributed in the interval $[0,1]$ (the second weight coefficient $\mu_2=1-\mu_1$ due to (\ref{eq:06})). Both the methods used the same parameters $r=2$ and $\varepsilon=0.01$ and the parameter $\beta=0.01$ was taken in AGS-SP. Each evolutionary method was launched 100 times and the HV index was averaged.
Results of the first experiment are presented in Table \ref{tab:01}.

\begin{table}[ht]
\caption{Results of solving the MMCO problem (\ref{eq:28}).}
\label{tab:01}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Algorithm} & \textbf{Number of trials} & \textbf{Number of Pareto points} & \textbf{HV}    \\ \hline
Pareto HV          &                           &                                  & \textbf{1.202} \\
AGS-R              & 300                       & 33                               & \textbf{1.146} \\
AGS-SP             & 226                       & 34                               & \textbf{1.146} \\
SMPSO              & 300                       & 15.75                            & 1.072          \\
OMOPSO             & 300                       & 15.89                            & 1.136          \\
NSGAII             & 300                       & 37.62                            & 0.913          \\
SPEA2              & 300                       & 38.32                            & 0.894          \\
IBEA               & 300                       & 41.99                            & 0.985          \\ \hline
\end{tabular}
\end{table}

The data of the table show that AGS-SP and AGS-R solve the problem (\ref{eq:28}) more efficient than evolutionary methods because while spending the same number of trials (and even less as AGS-SP) they provide the  better HV index and find more points in Pareto set approximation. If to compare AGS-SP and AGS-R with each other, then AGS-SP has an advantage due to HV indices being the same, but AGS-SP was "faster" as it executed fewer trials.

The second series was aimed to studying the influence of the parameter $\beta$ from (\ref{eq:24}) on number of trials produced by AGS-SP and HV index. Bi-criteria problems were considered where complicated multiextremal two-dimensional random functions \cite{Grishagin2018} were taken as partial criteria. The level curves of some of them are shown in Fig.  \ref{fig:2}. 50 MMCO problems of this type have been solved with the same parameters $r=3, \varepsilon=0.01$ and number $m=50$ of weight coefficient vectors (\ref{eq:11}) for different values of the parameter $\beta$ from the interval $[0,0.1]$. After solving 50 MMCO problems the results were averaged for each separate $\beta$. The change of HV and the number of trials depending on $\beta$ is shown in Fig. \ref{fig:3}.

As it can be seen, the growth of the parameter $\beta$ leads to significant decrease in the number of trials. At the same time, when increasing $\beta$, the value of the HV index deteriorates. As it follows from the graphs, the value $\beta=0.01$ can be taken as a good enough compromise.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{fig2}
\caption{The level curves for test function.} \label{fig:2}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{fig3}
\caption{Influence of the parameter $\beta$.} \label{fig:3}
\end{figure}

Problems with greater dimension were solved in the third series of experiments. Multiextremal functions of the class GKLS \cite{Gaviano2003} have been chosen as partial criteria of the MMCO problems.  In global optimization this class is widely used for testing global search algorithms. Bi-criteria problems of dimensions $n=3$ and $n=4$ were considered. Weight coefficients were chosen as nodes of a uniform grid in the domain defined by the conditions (\ref{eq:06}) with number of nodes equal to 200 in 3-dimensional case and to 50 for $n=4$. 100 MMCO problems were participated in the experiment for each dimension. ALR-R and ALR-SP used the same parameter   $r=4.5$.  ALR-SP applied the parameter $\beta=0.05$ for the dimension $n=3$ and $\beta=0.025$ in the case of $n=4$.  Running of each evolutionary method was repeated 10 times and results were averaged.

All the competing methods solved the set of 100 MMCO problems several times such that in each case they executed different number of trials and, correspondingly, obtained different values of HV's. If we take a method, consider all the pairs $(K, HV(K))$ where $K$ is number of trials executed by the method and $HV(K)$ is the corresponding value of the HV-index, and place all these pairs as points on the plane with coordinate abscissa axis $K$ and ordinate axis $HV$ we build so called \textit{operational characteristic} \cite{Grishagin2016_2} of the method. On the base of operational characteristics it is easy to compare visually the efficiency of different methods. Indeed, if the operational characteristic of one method is located above the characteristic of the other, then the first method is better because it spends fewer trials for obtaining the same HV, or, equivalently, the first method provides better HV for the same number of trials. Operational characteristics of the competing algorithms for dimensions 3 and 4 are presented in Fig. \ref{fig:4} and \ref{fig:5} correspondingly.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig4}
\caption{Operational characteristics for dimension 3.} \label{fig:4}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig5}
\caption{Operational characteristics for dimension 4.} \label{fig:5}
\end{figure}

As it follows from the results for the multidimensional cases, the undisputed winner is the proposed algorithm AGS-SP, i.e., the applied modifications enhance significantly the efficiency of solving complicated multiextremal multicriterial problems and allow obtaining better approximation of the Pareto set. And what's more, AGS-SP achieves the best possible value of HV-index in the 4-dimensional experiment while the other methods are far from the best level.

\section{Concludes}\label{sec:5}

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\bibliographystyle{splncs04}
\bibliography{bibliography}

% \begin{thebibliography}{8}
% \bibitem{ref_article1}
% Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

% \bibitem{ref_lncs1}
% Author, F., Author, S.: Title of a proceedings paper. In: Editor,
% F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
% Springer, Heidelberg (2016). \doi{10.10007/1234567890}

% \bibitem{ref_book1}
% Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
% Location (1999)

% \bibitem{ref_proc1}
% Author, A.-B.: Contribution title. In: 9th International Proceedings
% on Proceedings, pp. 1--2. Publisher, Location (2010)

% \bibitem{ref_url1}
% LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
% Oct 2017
% \end{thebibliography}
\end{document}

### 2023-06-10 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
chqck the last part of the paper

@@
 In general case, the functions satisfying the Lipschitz condition are multiextremal, therefore, the convolution (\ref{eq:05}) will be multiextremal and Lipschitzian as well.
-It means that for solving the problem (\ref{eq:08}) it is required to apply methods of non-linear programming which provides finding the global optimum.
+It means that for solving the problem (\ref{eq:08}) it is required to apply methods of nonlinear programming which provides finding the global optimum.
 There are many global optimization algorithms designed under different assumptions about properties of objective function; relevant references can be found, for instance, in monographs \cite{Pardalos2017,Marler2009,Sergeyev2013}.
@@
 The information-statistical algorithms \cite{Gergel2017_1,Gergel2018,Barkalov2014,Barkalov2015,Barkalov2018,Strongin2003}  as deterministic methods with guaranteed convergence are among the most effective for analyzing multiextremal optimization problems.
-For solving multidimensional problems, ISA algorithms apply two approaches based on ideas of dimensionality reduction.
+For solving multidimensional problems, these algorithms apply two approaches based on ideas of dimensionality reduction.
 In the framework of one approach the multidimensional problem is reduced to a family of univariate problems connected recursively (scheme of nested optimization) \cite{Grishagin2016,Zilinskas2015,Gergel2019_2}.
-The other approach \cite{Gergel2018,Barkalov2014,Sergeyev2013} exploits the known fact that $n$-dimensional box (1) and the closed interval $[0,1]$ of the real axis are equipotent sets and there exist unambiguous and continuous mappings of this interval onto the hyperparallelepiped (1) called \textit{Peano-type} curves or \textit{evolvents}.
+The other approach \cite{Gergel2018,Barkalov2014,Sergeyev2013} exploits the known fact that $n$-dimensional box (1) and the closed interval $[0,1]$ of the real axis are equipotent sets and there exist unambiguous and continuous mappings of this interval onto the hyperparallelepiped (1) called \textit{Peano-type curves} or \textit{evolvents}.
 
@@
 \begin{equation}
-    \min_{y\in H} g(y) = \min_{x\in[0,1]} g(y(x))
+    \min_{y\in H} g(y) = \min_{x\in[0,1]} g(y(x)),
 \end{equation}
-i.e., instead of minimization of the multidimensional function $g(y)$ in the domain (\ref{eq:01}) one can solve the minimization problem for the one-dimensional function $g(y(x))$ over interval $[0,1]$.
+i.e.
+instead of minimization of the multidimensional function $g(y)$ in the domain (\ref{eq:01}) one can solve the minimization problem for the one-dimensional function $g(y(x))$ over the interval $[0,1]$.
 
@@
     \label{eq:10}
-    |g(y(x')) - g(y(x''))| \leq G \|x'-x''\|, x', x'' \in [0,1]
+    |g(y(x')) - g(y(x''))| \leq G \left|x'-x''\right|^{1/n}, \; x', x'' \in [0,1],
 \end{equation}
@@
     \label{eq:11}
-    M=(\mu^1, \dots, \mu^m)
+    M=(\mu^1, \dots, \mu^m),
 \end{equation}
@@
     \label{eq:12}
-    \mu^i=(\mu^i_k, \dots, \mu^i_k), 1 \leq i \leq m.
+    \mu^i=(\mu^i_k, \dots, \mu^i_k), \; 1 \leq i \leq m.
 \end{equation}
@@
     \label{eq:13}
-    W(\mu^i, y)= V(\mu^i, y) + \alpha \sum^k_{s=1}\mu^i_s \varphi_s (y) \to \min, y \in H, 1 \leq i \leq m.
+    W(\mu^i, y)= V(\mu^i, y) + \alpha \sum^k_{s=1}\mu^i_s \varphi_s (y) \to \min, \; y \in H, 1 \leq i \leq m.
 \end{equation}
@@
     \label{eq:14}
-    f^i(x) = W(\mu^i, y(x)), 1 \leq i \leq m.
+    f^i(x) = W(\mu^i, y(x)), \; 1 \leq i \leq m.
 \end{equation}
@@
     \label{eq:15}
-    f^i(x) \to min, x \in [0,1], 1 \leq i \leq m,
+    f^i(x) \to min, \; x \in [0,1], \; 1 \leq i \leq m,
 \end{equation}
@@
 The following computational scheme is proposed for solving the problems of the family (\ref{eq:15}).
-In this scheme, all the problems (\ref{eq:15}) are solved sequentially and during minimization of current function $f^i(x), i > 1$, the results of preceding problems $f^j(x) \to \min, 1 \leq j \leq i - 1$ completed already are taken into account.
+In this scheme, all the problems (\ref{eq:15}) are solved sequentially and during minimization of current function $f^i(x), i > 1$, the results of preceding problems $f^j(x) \to \min, 1 \leq j \leq i - 1,$ completed already are taken into account.
 
@@
 
-Hereinafter the term ``trial'' will denote the evaluation of the objective function $f^i(x)$ at a point $x^j \in [0,1]$ and obtaining the vector (result of trial)
+Hereinafter the term ``trial'' will denote the evaluation of the objective function $f^i(x)$ at a point $x^j \in [0,1]$ and obtaining the vector (trial result)
 \begin{equation}
@@
 Before starting the described algorithm to build a Pareto set approximation, the set (\ref{eq:11}) of weight coefficient vectors is supposed to be given.
-In the framework of the statement (\ref{eq:01})-(\ref{eq:03}) this set generates the family of $m$ multidimensional scalar problems (\ref{eq:13}) and corresponding to those the family of $m$ univariate problems (\ref{eq:13}).
+In the framework of the statement (\ref{eq:01})--(\ref{eq:03}) this set generates the family of $m$ multidimensional scalar problems (\ref{eq:13}) and corresponding to those the family of $m$ univariate problems (\ref{eq:15}).
 
@@
 At the first stage when $f^i(x)=f^1(x)$, $q=2$ and $x^1=0, x^2=1$.
-These 2 trials generate results $\Omega^1(x^1), \Omega^1(x^2)$ from (\ref{eq:15a}) and initial information matrix $A^1_2$ from (\ref{eq:18}).
+These two trials generate results $\Omega^1(x^1), \Omega^1(x^2)$ from (\ref{eq:15a}) and initial information matrix $A^1_2$ from (\ref{eq:18}).
 
@@
 \end{equation}
-
-The values $u_j = f^i(x_j), 1\leq j \leq p$ are juxtaposed to the points $x_j, 1\leq j \leq p$.
+The values $u_j = f^i(x_j), 1\leq j \leq p,$ are juxtaposed to the points $x_j, 1\leq j \leq p$.
 
@@
   \begin{matrix}
-     \gamma \Gamma & \Gamma >0 \\
-     1 & \Gamma = 0 
+     \gamma \Gamma, & \Gamma >0 \\
+     1, & \Gamma = 0 
   \end{matrix}
@@
 \textbf{Rule 7.} Neighboring points from (\ref{eq:19}) provide a partitioning of the interval $[0, 1]$ into $p-1$ subintervals $(x_{j-1}, x_j), 1 \leq j \leq p-1$.
-
 A real value $B(j)$ is assigned to each subinterval $(x_{j-1}, x_j), 1 \leq j \leq p-1$. $B(j)$ is called \textit{the characteristic}  of this subinterval and is calculated as
@@
     \label{eq:24}
-    B(j) = \rho_j + \frac{(u_j-u_{j-1})^2}{\gamma^2 \rho_j} - \frac{2 (u_j+u_{j-1})}{\gamma} + \beta (d'_j + d'_{j-1})
+    B(j) = \rho_j + \frac{(u_j-u_{j-1})^2}{\gamma^2 \rho_j} - \frac{2 (u_j+u_{j-1})}{\gamma} + \beta (d'_j + d'_{j-1}),
 \end{equation}
@@
 \end{equation}
-is true where $\varepsilon > 0$ is predefined accuracy, then the $i$-th  stage (minimization of function $f^i(x)$) is over.
+is true, where $\varepsilon > 0$ is predefined accuracy, then the $i$-th  stage (minimization of function $f^i(x)$) is over.
 
@@
 To obtain the executable program code the compilers Intel C++ 17.0.0 and python 3.9 were used.
-Computations were performed with the support of the package scikit-learn 0.24.2 and the applied optimization system Globalizer \cite{globalizerSystem}.
+Computations were performed with the support of the package scikit-learn 0.24.2 and the optimization software Globalizer \cite{globalizerSystem}.
 
 The experiment was organized in the following way.
-For efficiency evaluation of the proposed algorithm hereinafter called AGS-SP (Algorithm of Global Search with Separating Planes), it was compared with its modification AGS-R (Algorithm of Global Search – Reduced) where computational scheme is similar, but Rules 3- 6 are removed and in the characteristic (\ref{eq:24}) parameter $\beta = 0$, AGS-R ignores the information about the Pareto set approximation.
-Moreover, to expand the basis of the comparison, some known meta-heuristic evolutionary MMCO methods were involved into the experiment, namely, algorithms SMPSO \cite{NDG09}, OMOPSO \cite{RC05}, NSGA-II \cite{DPA02}, SPEA2 \cite{ZLT01}, IBEA \cite{ZK04} from the library jMetal \cite{DNA10}.
+For efficiency evaluation of the proposed algorithm hereinafter called AGS-SP (Algorithm of Global Search with Separating Planes), it was compared with its modification AGS-R (Algorithm of Global Search Reduced) where computational scheme is similar, but Rules 3 -- 6 are removed and in the characteristic (\ref{eq:24}) parameter $\beta = 0$.
+AGS-R ignores the information about the Pareto set approximation.
+Moreover, to expand the basis of the comparison, some known meta-heuristic evolutionary MMCO methods were involved into the experiment, namely, algorithms SMPSO \cite{NDG09}, OMOPSO \cite{RC05}, NSGA-II \cite{DPA02}, SPEA2 \cite{ZLT01}, IBEA \cite{ZK04} from the jMetal framework \cite{DNA10}.
 
@@
 The first one dealt with a simple problem and was aimed at demonstration of positive influence of criteria space analysis by means of separating planes and at comparing with evolutionary techniques.
-
 This problem was bi-criteria one \cite{CHIANDUSSI2012912}:
@@
     \begin{matrix}
-f_1 (y)=y_1 \\
-f_2 (y)=(1+10y_2 )[1- \left(\frac{y_1}{1+10y_2} \right)^2- \frac{y_1}{1+10y_2} \sin(8 \pi y_1 ) ] \\
-0 \leq y_1,y_2 \leq 1
+f_1 (y)=y_1, \\
+f_2 (y)=(1+10y_2 )[1- \left(\frac{y_1}{1+10y_2} \right)^2- \frac{y_1}{1+10y_2} \sin(8 \pi y_1 ) ], \\
+0 \leq y_1,y_2 \leq 1.
     \end{matrix}
@@
 \label{tab:01}
+\center
 \begin{tabular}{|c|c|c|c|}
@@
 \textbf{Algorithm} & \textbf{Number of trials} & \textbf{Number of Pareto points} & \textbf{HV}    \\ \hline
-Pareto HV          &                           &                                  & \textbf{1.202} \\
+Pareto HV          & ---                       & ---                              & \textbf{1.202} \\
 AGS-R              & 300                       & 33                               & \textbf{1.146} \\
@@
 The data of the table show that AGS-SP and AGS-R solve the problem (\ref{eq:28}) more efficient than evolutionary methods because while spending the same number of trials (and even less as AGS-SP) they provide the  better HV index and find more points in Pareto set approximation.
-If to compare AGS-SP and AGS-R with each other, then AGS-SP has an advantage due to HV indices being the same, but AGS-SP was "faster" as it executed fewer trials.
+If to compare AGS-SP and AGS-R with each other, then AGS-SP has an advantage due to HV indices being the same, but AGS-SP was ``faster'' as it executed fewer trials.
 

### 2023-06-10 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
1. Заключение 2. Сокрашение самоцитирования

@@
 
-In the class of deterministic methods for multiextremal optimization with guaranteed convergence to global optimum \cite{Gablonsky2001,Sergeyev2010,Sergeyev2006,Evtushenko2009,Evtushenko2014,Barkalov2015,Strongin2003,Sergeyev2013}, the algorithms developed in the framework of the information-statistical approach (ISA) \cite{Gergel2017_1,Gergel2018,Barkalov2014,Barkalov2018} are among the most effective.
+In the class of deterministic methods for multiextremal optimization with guaranteed convergence to global optimum \cite{Gablonsky2001,Sergeyev2010,Sergeyev2006,Evtushenko2014,Strongin2003,Sergeyev2013}, the algorithms developed in the framework of the information-statistical approach (ISA) \cite{Gergel2018,Barkalov2018} are among the most effective.
 The present paper describes a new algorithm belonging to the ISA class and aimed at solving MMCO problems that accelerates the Pareto set building by the application of machine learning ideas.
@@
 Following this way, the initial MMCO problem is reduced to a family of single-criterion optimization subproblems by means of parameterized convolution.
-Then each multidimensional subproblem is transformed to an equivalent univariate (one-dimensional) optimization problem formed on the base of Peano-type mapping \cite{Evtushenko2014,Gergel2017_1,Gergel2018,Marler2004,Pardalos2017}.
+Then each multidimensional subproblem is transformed to an equivalent univariate (one-dimensional) optimization problem formed on the base of Peano-type mapping \cite{Evtushenko2014,Gergel2018,Marler2004,Pardalos2017}.
 Further, a special global optimization algorithm designed on the ISA platform is applied for solving the univariate problems.
@@
 In our research we took weight coefficients as nodes of a uniform grid in the polyhedron (\ref{eq:06}).
-For assessment of the quality of Pareto set approximation obtained after solving problems (\ref{eq:08}) with mentioned weights, the HyperVolume index (HV-index) \cite{Evtushenko2014,Gergel2017_1,Gergel2018} was applied as traditional criterion of approximation accuracy.
+For assessment of the quality of Pareto set approximation obtained after solving problems (\ref{eq:08}) with mentioned weights, the HyperVolume index (HV-index) \cite{Evtushenko2014,Gergel2018} was applied as traditional criterion of approximation accuracy.
 
@@
 
-The information-statistical algorithms \cite{Gergel2017_1,Gergel2018,Barkalov2014,Barkalov2015,Barkalov2018,Strongin2003}  as deterministic methods with guaranteed convergence are among the most effective for analyzing multiextremal optimization problems.
+The information-statistical algorithms \cite{Gergel2018,Barkalov2018,Strongin2003}  as deterministic methods with guaranteed convergence are among the most effective for analyzing multiextremal optimization problems.
 For solving multidimensional problems, these algorithms apply two approaches based on ideas of dimensionality reduction.
-In the framework of one approach the multidimensional problem is reduced to a family of univariate problems connected recursively (scheme of nested optimization) \cite{Grishagin2016,Zilinskas2015,Gergel2019_2}.
-The other approach \cite{Gergel2018,Barkalov2014,Sergeyev2013} exploits the known fact that $n$-dimensional box (1) and the closed interval $[0,1]$ of the real axis are equipotent sets and there exist unambiguous and continuous mappings of this interval onto the hyperparallelepiped (1) called \textit{Peano-type curves} or \textit{evolvents}.
+In the framework of one approach the multidimensional problem is reduced to a family of univariate problems connected recursively (scheme of nested optimization) \cite{Gergel2019_2}.
+The other approach \cite{Gergel2018,Sergeyev2013} exploits the known fact that $n$-dimensional box (1) and the closed interval $[0,1]$ of the real axis are equipotent sets and there exist unambiguous and continuous mappings of this interval onto the hyperparallelepiped (1) called \textit{Peano-type curves} or \textit{evolvents}.
 
@@
 
-\section{Concludes}\label{sec:5}
+\section{Conclusion}\label{sec:5}
+
+In the paper a new numerical method for solving multidimensional multicriterial  optimization problems has been proposed for the case of multiextremal black-box partial criteria.
+The novelty of the method consists in application of logistic regression as a tool, widely used in machine learning, for planning iterations of the global optimization algorithm.
+The workability and efficiency of the method have been evaluated in a representative experiment on test classes of multiextremal problems of different dimensions.
+In the experiment the proposed method was compared to its version which does not utilize logistic regression and to several known evolutionary methods.
+All the experiments have demonstrated significant advantages of the new method over other competitors.
+
+The further development of the research will be associated with designing parallel generalizations of the method exploiting ideas of parallel characteristical algorithms and oriented at high-performance  supercomputers.
 

### 2023-06-10 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
reduce the size of the figures
link replacement

@@
 Computations were conducted on the supercomputer ``Lobachevsky'' of the Nizhni Novgorod State University (operating system CentOS 7, control system SLURM).
-One node of the supercomputer has 2 processors Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM.
+%One node of the supercomputer has 2 processors Intel Sandy Bridge E5-2660 2.2 GHz, 64 Gb RAM.
 One CPU contains 8 cores, i.e., 16 cores are available on one node.
@@
 The second series was aimed to studying the influence of the parameter $\beta$ from (\ref{eq:24}) on number of trials produced by AGS-SP and HV index.
-Bi-criteria problems were considered where complicated multiextremal two-dimensional random functions \cite{Grishagin2018} were taken as partial criteria.
+Bi-criteria problems were considered where complicated multiextremal two-dimensional random functions \cite{Grishagin2015_2} were taken as partial criteria.
 The level curves of some of them are shown in Fig.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{fig3}
+\includegraphics[width=0.7\textwidth]{fig3}
 \caption{Influence of the parameter $\beta$.} \label{fig:3}
@@
 \centering
-\includegraphics[width=\textwidth]{fig4}
+\includegraphics[width=0.9\textwidth]{fig4}
 \caption{Operational characteristics for dimension 3.} \label{fig:4}
@@
 \centering
-\includegraphics[width=\textwidth]{fig5}
+\includegraphics[width=0.9\textwidth]{fig5}
 \caption{Operational characteristics for dimension 4.} \label{fig:5}

### 2023-06-10 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
fix

@@
 As a core of novelty, this algorithm includes in its computational scheme a machine learning procedure with combination of accumulating the information of solved subproblems that allows one to significantly accelerate building the Pareto set.
-Effectiveness of the proposed approach is estimated in the representative computational experiment on test sets of multicriterial problems with multiextremal criteria for different dimensions and number of criteria in comparison with several nature-inspired multiobjective optimization algorithms.
+Effectiveness of the proposed approach is estimated in the representative computational experiment on test sets of multicriterial problems with multiextremal criteria for different dimensions in comparison with several nature-inspired multiobjective optimization algorithms.
 

### 2023-06-10 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
fix

@@
 Artificial intelligence models formulated as decision-making problems are extremely diverse and attract significant attention of researchers due to the richness of theoretical background and practical importance.
-The high complexity of decision-making models, along with the use of existing effective approaches to their analysis, requires the development of new ideas that overcome these difficulties.
+The high complexity of decision-making models requires the development of new ideas that reduce this complexity.
 A promising way to improve decision-making algorithms is to combine in their development the advantages of different approaches.
@@
 
-For efficiency confirmation of the proposed approach combining the global optimization algorithm with machine learning procedures, the representative computational experiment has been conducted on the test class of multiextremal multicriterial problems of different dimensions and number of criteria.
+For efficiency confirmation of the proposed approach combining the global optimization algorithm with machine learning procedures, the representative computational experiment has been conducted on the test class of multiextremal multicriterial problems of different dimensions.
 For expanding the confirmation area, the test problems were solved by several known meta-heuristic methods of multicriterial optimization.

### 2023-06-10 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
fix

@@
 In our research we took weight coefficients as nodes of a uniform grid in the polyhedron (\ref{eq:06}).
-For assessment of the quality of Pareto set approximation obtained after solving problems (\ref{eq:08}) with mentioned weights, the HyperVolume index (HV-index) \cite{Evtushenko2014,Gergel2018} was applied as traditional criterion of approximation accuracy.
+For assessment of the quality of Pareto set approximation obtained after solving problems (\ref{eq:08}) with mentioned weights, the HyperVolume index (HV-index) \cite{Evtushenko2014,Gergel2018} was applied as a traditional criterion of the approximation accuracy.
 
@@
 In general case, the functions satisfying the Lipschitz condition are multiextremal, therefore, the convolution (\ref{eq:05}) will be multiextremal and Lipschitzian as well.
-It means that for solving the problem (\ref{eq:08}) it is required to apply methods of nonlinear programming which provides finding the global optimum.
+It means that for solving the problem (\ref{eq:08}) it is required to apply methods of nonlinear programming which provide finding the global optimum.
 There are many global optimization algorithms designed under different assumptions about properties of objective function; relevant references can be found, for instance, in monographs \cite{Pardalos2017,Marler2009,Sergeyev2013}.
@@
 
-Existence of Peano mappings can be applied to solving the optimization problems in the following way.
+Peano mappings can be applied to solving the optimization problems in the following way.
 Let there be a $n$-dimensional continuous function $g(y)$ defined in the domain (\ref{eq:01}) and $y(x)$ is a Peano-type curve.
@@
 \end{equation}
-after implementing $p$ trials at points $x^1, x^2, \dots, x^p$ when minimizing a function $f'(x)$ from (\ref{eq:15}).
+after implementing $p$ trials at points $x^1, x^2, \dots, x^p$ when minimizing a function $f^i(x)$ from (\ref{eq:15}).
 
@@
 
-\textbf{Rule 3.} In the set of trial coordinates $x_j, 1 \leq j \leq p$, non-dominated points are selected according the rule of Definition 2 and all such points are considered as points belonging to current approximation of the Pareto set.
+\textbf{Rule 3.} In the set of trial coordinates $x_j, 1 \leq j \leq p$, non-dominated points are selected according to the rule of Definition 2 and all such points are considered as points belonging to current approximation of the Pareto set.
 All columns of the information matrix $A^i_p$ are marked up with labels 0 or 1 where 1 corresponds to non-dominated points and 0 to dominated ones.
@@
 
-\textbf{Rule 8.} Among all subintervals, the subinterval $(x_{j-1}, x_j), 1 \leq j \leq p-1$, with the highest characteristic is chosen:
+\textbf{Rule 8.} Among all subintervals, the subinterval $(x_{l-1}, x_l), 1 \leq l \leq p-1$, with the highest characteristic is chosen:
 \begin{equation}
@@
 \centering
-\includegraphics[width=0.8\textwidth]{fig2}
+\includegraphics[width=\textwidth]{fig2}
 \caption{The level curves for test function.} \label{fig:2}

### 2023-06-10 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
fix

@@
 100 MMCO problems were participated in the experiment for each dimension.
-ALR-R and ALR-SP used the same parameter   $r=4.5$.
-ALR-SP applied the parameter $\beta=0.05$ for the dimension $n=3$ and $\beta=0.025$ in the case of $n=4$.
+AGS-R and AGS-SP used the same parameter $r=4.5$.
+AGS-SP applied the parameter $\beta=0.05$ for the dimension $n=3$ and $\beta=0.025$ in the case of $n=4$.
 Running of each evolutionary method was repeated 10 times and results were averaged.

### 2023-06-13 2023/ITMO paper/elsarticle-template-num.tex
ITMO paper template

%% 
%% Copyright 2007-2020 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%%
%% 
%%
%% $Id: elsarticle-template-num.tex 190 2020-11-23 11:12:32Z rishi $
%%
%%
\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
\usepackage{lineno}

\journal{Knowledge-Based Systems}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%% \fntext[label3]{}

\title{Title of the paper}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

\author[UNN,ITMO]{Konstantin Barkalov}
\author[UNN,ITMO]{Denis Rodionov}

\affiliation[UNN]{organization={Lobachevsky University},%Department and Organization
            addressline={Gagarin Av. 23}, 
            city={Nizhni Novgorod},
            postcode={603022}, 
            %state={},
            country={Russia}}

\affiliation[ITMO]{organization={ITMO University},%Department and Organization
            addressline={Lomonosova St. 9}, 
            city={St. Petersburg},
            postcode={191002}, 
            %state={},
            country={Russia}}

\begin{abstract}
%% Text of abstract

\end{abstract}

%%Graphical abstract
%\begin{graphicalabstract}
%\includegraphics{grabs}
%\end{graphicalabstract}

%%Research highlights
\begin{highlights}
\item Research highlight 1
\item Research highlight 2
\end{highlights}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

\linenumbers

%% main text
\section{Introduction}
\label{sec_intro}

\section{Related work}
\label{sec_rel}

\section{Lipschitz optimization} 
\label{sec_lip}
Test reference \cite{Strongin2000,Barkalov2021}.

\section{iOpt framework} 
\label{sec_iOpt}

\section{Numerical experiments}
\label{sec_exp}

\section*{CRediT authorship contribution statement}

Konstantin Barkalov:
Denis Rodionov:

\section*{Declaration of competing interest}

The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

\section*{Acknowledgements}

This research was supported by the research center ``Strong Artificial Intelligence in Industry'' of ITMO University.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
\bibliographystyle{elsarticle-num} 
\bibliography{bibliography.bib}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

%\begin{thebibliography}{00}

%% \bibitem{label}
%% Text of bibliographic item

%\bibitem{}

%\end{thebibliography}
\end{document}
\endinput
%%
%% End of file `elsarticle-template-num.tex'.

### 2023-06-13 2023/ITMO paper/elsarticle-template-num.tex
Draft ot the Global Optimization section

@@
 %%
-\documentclass[preprint,12pt]{elsarticle}
+%\documentclass[preprint,12pt]{elsarticle}
 
 %% Use the option review to obtain double line spacing
-%% \documentclass[preprint,review,12pt]{elsarticle}
+ \documentclass[preprint,review,12pt]{elsarticle}
 
@@
 %% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
+%%\usepackage{amsthm}
 
@@
 \usepackage{lineno}
+
+%% Русский язык - закомментировать в финальной версии
+\usepackage[russian]{babel}
+
+%% advanced mathematics
+\usepackage{amsmath}
 
@@
 \label{sec_intro}
+Подбор гиперпараметров - важная и сложная задача.
+
+Много разных фреймворков, у всех есть свои достоинства и недостатки.
 
@@
 
+Простые задачи - метод перебора
+
+Задачи большой размерности - метаэвристические алгоритмы
+
+Задачи малой размерности - байесовская оптимизация.
+
+Липшицева оптимизация - основные понятия, ссылки на работы. Соотношение с другими подходами. Проблема дискртетных параметров.
+
 \section{Lipschitz optimization} 
 \label{sec_lip}
+
+\subsection{Problem statement}
+
+Мы будем рассматривать задачу поиска глобального минимума в следующей постановке:
+\begin{gather}
+	\varphi(y^*) = \min \varphi(y), \; y \in D, \label{f_func} \\
+	D = \left\{y \in R^N : a_j \leq y_j \leq b_j , \; 1 \leq j \leq N \right\}. \label{f_D}
+\end{gather}
+
+Будем также предполагать, что целевая функция $\varphi(y)$ удовлетворяет в области поиска $D$ условию Липшица, т.е.
+\begin{equation} \label{f_lip}
+	\left| \varphi(y')-\varphi(y'') \right| \leq L\left\| y' - y''  \right\| , \; y',y'' \in D, \; 0<L<\infty.
+\end{equation}
+Здесь $ \left\| \cdot \right\|$ обозначает евклидову норму, а константа $L$ является априори не изветной и подлежит оценке в процессе решения. 
+Такие задачи называются задачами липшицевой глобальной оптимизации with box-constraints.
+
+Функция $\varphi(y)$ предполагается многоэкстремальной и заданной в виде ``черного ящика'', т.е.
+некоторого алгоритма вычисления ее значений.
+Кроме того, подразумевается, что каждое поисковое испытание (т.е.
+вычисление значения функции в точке допустимой области) требует значительных вычислительных ресурсов.
+Такая постановка проблемы полностью соответствует задаче настройки гиперпараметров методов машинного обучения (see ).
+
+Условие Липшица допускает следующую геометрическую интерпретацию.
+Допустим, что одномерная липшицева функция $f(x)$ (с известной константой Липшица $L$) была вычислена в двух точках $x'$ и $x''$.
+В соответствии с условием (\ref{f_lip}) выполняются следующие неравенства, характеризующие поведение функции $f(x)$ на интервале $[x', x'']$:
+\begin{gather*}
+	f(x) \leq f(x') + L(x-x'), \; x \geq x',\\
+	f(x) \geq f(x') - L(x-x'), \; x \geq x',\\
+	f(x) \leq f(x') - L(x-x''), \; x \leq x'',\\
+	f(x) \geq f(x') + L(x-x''), \; x \leq x''.
+\end{gather*}
+
+В силу данных неравенств значения функции в точках интервала $[x', x'']$ должны располагаться внутри области, ограниченной прямыми, проходящими через точки $(x', f(x'))$ и $(x'', f(x''))$ с наклоном $L$ и $-L$ (см.
+рис.
+).
+
+Также в соответствии с (\ref{f_lip}) может быть записано неравенство $f(x) \geq F(x)$, где функция $F(x)$ (называемая минорантой) определяется в соответствии с формулой
+\[
+F(x) = \max\left\{f(x') - L(x-x'),f(x') + L(x-x'')\right\}, \; x\in [x', x''].
+\]
+
+Наименьшее значение миноранты $F(x)$ на $[x', x'']$ совпадает с оценкой наименьшего значения функции $f(x)$ на данном интервале. Указанная оценка достигается в точке 
+\[
+\overline{x} = \frac{x'+x''}{2}-\frac{f(x'')-f(x')}{2L}
+\] 
+и равна
+\[
+F(\overline{x}) = \frac{f(x')+f(x'')}{2} -L \frac{x''-x'}{2}.
+\]
+
+Методы липшицевой глобальной оптимизации используют эту идею в своих вычислительных правилах; на ней также основано и доказательство их сходимости (see, e.g., ).
+
+\subsection{Проблема редукции размерности}
+
+Одной из основных трудностей при решении многомерных задач глобальной оптимизации является рост вычислительных затрат при увеличении размерности задачи.
+Эта сложность имеет место и при решении задач липшицевой глобальной оптимизации.
+
+Уменьшение числа испытаний при тех же требованиях к точности решения возможно за счет более полного использования априорных предположений о минимизируемой функции, что приводит к адаптивным последовательным алгоритмам оптимизации.
+
+Известным подходом к решению многомерной задачи (\ref{f_func}) является ее редукция к одномерной задаче с последующим применением эффективных одномерных алгоритмов глобальной оптимизации (see).
+Здесь хорошо зарекомендовал себя подход, в котором размерность задачи редуцируется при помощи кривой Пеано–Гильберта $y(x), x \in [0, 1]$.
+
+Такая кривая однозначно и непрерывно отображает интервал $[0, 1]$ на гиперинтервал $D$ из (\ref{f_D}), т.е. заполняет область $D$.
+
+Для решения задачи глобальной липшицевой оптимизации кривые, заполняющие пространство, могут быть применены следующим образом.
+Если $y(x)$ есть кривая Пеано–Гильберта, то из непрерывности целевой функции $\varphi(y)$ следует, что
+\[
+\min
+\]
+т.е. исходная многомерная задача (\ref{f_func}) редуцируется к одномерной.
+
+При этом известно (see []), что если многомерная функция $\varphi(y), ; y \in D,$  удовлетворяет условию Липшица с константой $L$, то редуцированная одномерная функция $\varphi(y(x)), ; x \in [0,1]$ удовлетворяет условию Гельдера
+\[
+holder
+\]
+Здесь $N$ есть размерность исходной задачи, а коэффициент определяется как .
+
+Замечание 1. Для редуцированной одномерной функции $\varphi(y(x))$  не будет выполняться условие Липшица. Однако многие одномерные алгоритмы липшицевой оптимизации
+могут быть обобщены на случай минимизации гeльдеровых функций (see []). Конкретный пример такого обобщения приведен в subsection \ref{sec_GSA}.
+
+Замечание 2.
+Теоретическая кривая Пеано-Гильберта $y(x)$ определяется как предельный объект.
+В численных алгоритмах применяются развертки $y_m(x)$, аппроксимирующие истинную кривую Пеано–Гильберта c заданным уровнем точности $2^{-m}$, зависящим от требуемой точности поиска.
+Эффективные схемы вычисления нескольких видов разверток подробно описаны в [].
+
+\subsection{Проблема дискретных параметров}
+
+Проблема, обзор подходов.
+
+Рассмотрим теперь случай, когда целевая функция задачи зависит от двух векторов параметров: вектора $y$, принадлежащего гиперинтервалу $D$, и вектора $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е.
+\begin{gather}\label{problem_i}
+\min{\left\{ \varphi(y,u) : y\in D \right\}},\\
+D=\left\{a_j \leq y_j \leq b_j, \; 1\leq j \leq N \right\}.\nonumber
+\end{gather}
+
+Такие конечные наборы значений характеризуют множество возможных сочетаний категориальных гиперпараметров исследуемого алгоритма машинного обучения. 
+%Целочисленные параметры = непрерывные параметры с округлением, добавить комментарий об этом.
+
+Занумеруем целыми числами $s, 1\leq s \leq S,$ все различные комбинации категориальных параметров, т.е. сопоставим каждому номеру $s$ вектор $u_s$. 
+Тогда рассматриваемая задача может быть записана в виде 
+\begin{gather}\label{problem_is}
+ \min_{s\in\{1,...,S\}}\left[\min{\left\{ \varphi(y,u_s) : y\in D \right\}}\right].
+\end{gather}
+
+Используя схему редукции размерности с помощью кривых Пеано $y(x), x\in [0,1]$ можно сопоставить каждой задаче минимизации по $y$ одномерную задачу минимизации
+\[
+ \min{\left\{ \varphi(y(x),u_s): x \in [0,1] \right\}}, s \in \{1,...,S\}.
+\]
+
+Рассмотрим теперь отображение 
+\[
+Y(x)=y(x-E(x)), \; x\in[0,S],
+\]
+переводящее любую точку интервала [0,S] на область $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функцию 
+\[
+f(x) = \varphi(Y(x),u_{E(x)+1}), x\in[0,S],
+\]
+имеющую, вообще говоря, разрывы в целочисленных точках $x_i = i, 1\leq i \leq S-1$.
+Поэтому значения  $z_i = f(x_i)$ в этих точках будем считать неопределенным, а значения индекса -- равным 0, т.е. $\nu(x_i) = 0$.
+
+Используя введенные обозначения можно переформулировать исходную задачу как
+\begin{equation}\label{problem_is1}
+\min \left\{f(x): x \in [0,S] \right\}.
+\end{equation}
+
+Применяя к решению задачи (\ref{problem_is1}) алгоритм глобального поиска, мы найдем решение задачи (\ref{problem_i}).
+При этом основная часть испытаний будет проведена в $s$-й подзадаче, решение которой соответствует решению исходной задачи (\ref{problem_i}).
+В остальных подзадачах будет проведена лишь незначительная часть испытаний, т.к.
+решения данных подзадач являются локально-оптимальными по отношению к решению $s$-й подзадачи.
+
 Test reference \cite{Strongin2000,Barkalov2021}.
@@
 \label{sec_iOpt}
+
+\subsection{Global search algorithm}
+\label{sec_GSA}
+
+\subsection{GSA for the problems with categorical parameters}
+\label{sec_mGSA}
 

### 2023-06-13 2023/ITMO paper/elsarticle-template-num.tex
add text in section 2

@@
 %%
-%\documentclass[preprint,12pt]{elsarticle}
+\documentclass[preprint,12pt]{elsarticle}
 
 %% Use the option review to obtain double line spacing
- \documentclass[preprint,review,12pt]{elsarticle}
+%% \documentclass[preprint,review,12pt]{elsarticle}
 
@@
 Здесь $ \left\| \cdot \right\|$ обозначает евклидову норму, а константа $L$ является априори не изветной и подлежит оценке в процессе решения. 
-Такие задачи называются задачами липшицевой глобальной оптимизации with box-constraints.
+%Такие задачи называются задачами липшицевой глобальной оптимизации with box-constraints.
 
@@
 Эта сложность имеет место и при решении задач липшицевой глобальной оптимизации.
+Например, решение задачи (\ref{f_func}) c помощью поиска по равномерной сетке с шагом $\epsilon$ потребует проведения 
+\[
+K \approx \prod_{j=1}^N{\frac{b_j-a_j}{\epsilon}}
+\]
+испытаний. Очевидно, что $K$ экспоненциально возрастает с ростом $N$.
+Уменьшение числа испытаний при тех же требованиях к точности решения возможно за счет адаптивного построения неравномерной сетки в области изменения параметров.
 
-Уменьшение числа испытаний при тех же требованиях к точности решения возможно за счет более полного использования априорных предположений о минимизируемой функции, что приводит к адаптивным последовательным алгоритмам оптимизации.
-
-Известным подходом к решению многомерной задачи (\ref{f_func}) является ее редукция к одномерной задаче с последующим применением эффективных одномерных алгоритмов глобальной оптимизации (see).
+Известным подходом к решению многомерной задачи (\ref{f_func}) является ее редукция к одномерной задаче с последующим применением эффективных одномерных алгоритмов глобальной оптимизации (see []).
 Здесь хорошо зарекомендовал себя подход, в котором размерность задачи редуцируется при помощи кривой Пеано–Гильберта $y(x), x \in [0, 1]$.
-
-Такая кривая однозначно и непрерывно отображает интервал $[0, 1]$ на гиперинтервал $D$ из (\ref{f_D}), т.е. заполняет область $D$.
+Такая кривая однозначно и непрерывно отображает интервал $[0, 1]$ на гиперинтервал $D$ из (\ref{f_D}), т.е. заполняет всю область $D$.
 
@@
 \[
-\min
+\min_{y \in D } \varphi(y) = \min_{x \in [0,1] } \varphi(y(x)),
 \]
@@
 
-При этом известно (see []), что если многомерная функция $\varphi(y), ; y \in D,$  удовлетворяет условию Липшица с константой $L$, то редуцированная одномерная функция $\varphi(y(x)), ; x \in [0,1]$ удовлетворяет условию Гельдера
+При этом известно (see []), что если многомерная функция $\varphi(y), \; y \in D,$  удовлетворяет условию Липшица с константой $L$, то редуцированная одномерная функция $\varphi(y(x)), \; \in [0,1]$ удовлетворяет the H{\"o}lder condition
 \[
-holder
+\left|\varphi(y(x'))-\varphi(y(x''))\right|\leq H\left|x'-x''\right|^{1/N}, \; x',x''\in[0,1].
 \]
-Здесь $N$ есть размерность исходной задачи, а коэффициент определяется как .
+Здесь $N$ есть размерность исходной задачи, а коэффициент $ H=2 L \sqrt{N+3}$.
 
@@
 В численных алгоритмах применяются развертки $y_m(x)$, аппроксимирующие истинную кривую Пеано–Гильберта c заданным уровнем точности $2^{-m}$, зависящим от требуемой точности поиска.
-Эффективные схемы вычисления нескольких видов разверток подробно описаны в [].
+Эффективные схемы вычисления нескольких видов разверток подробно описаны в \cite{Sergeyev2013}.
+%пример развертки - рисунок с функцией
 
@@
 \bibliographystyle{elsarticle-num} 
-\bibliography{bibliography.bib}
+\bibliography{bibliography}
 

### 2023-06-14 2023/ITMO paper/elsarticle-template-num.tex
add some text in section 2

@@
 \usepackage{amsmath}
+\usepackage[]{algorithm2e}
+\usepackage{algorithmic}
 
@@
 
-Проблема, обзор подходов.
+Особый интерес представляют задачи, в которых часть параметров может принимать значения из заранее заданного конечного множества, т.к.
+для них сложнее построить оценки оптимума по сравнению с задачами с непрерывными параметрами.
 
-Рассмотрим теперь случай, когда целевая функция задачи зависит от двух векторов параметров: вектора $y$, принадлежащего гиперинтервалу $D$, и вектора $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е.
+Методам решения задач со смешанными параметрами посвящено много публикаций (см., например, обзоры \cite{Burer2012,Boukouvala2016}). 
+Известные детерминированные методы решения задач данного класса ориентированы на решение линейных или выпуклых задач.
+Многоэкстремальные задачи предлагается решать с помощью метаэвристических алгоритмов.
+
+Нами был предложен и реализован новый детерминированный подход к решению задач со смешанными параметрами; приведем здесь его краткое описание.
+
+Пусть  целевая функция задачи зависит от двух векторов параметров: вектора $y$, принадлежащего гиперинтервалу $D$, и вектора $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е.
 \begin{gather}\label{problem_i}
@@
 
-Такие конечные наборы значений характеризуют множество возможных сочетаний категориальных гиперпараметров исследуемого алгоритма машинного обучения. 
+Такие конечные наборы значений могут характеризовать множество возможных сочетаний категориальных гиперпараметров исследуемого алгоритма машинного обучения. 
 %Целочисленные параметры = непрерывные параметры с округлением, добавить комментарий об этом.
@@
 
+А нужно ли тут отдельное описание?
+
 \section{Numerical experiments}

### 2023-06-21 2023/ITMO paper/elsarticle-template-num.tex
some corrections
extend plan of the paper

@@
 \usepackage{amsmath}
-\usepackage[]{algorithm2e}
-\usepackage{algorithmic}
+\usepackage[ruled]{algorithm2e}
 
@@
 \author[UNN,ITMO]{Konstantin Barkalov}
+\author[UNN,ITMO]{Evgeny Kozinov}
+\author[UNN,ITMO]{Ilya Lebedev}
+\author[UNN,ITMO]{Denis Karchkov}
 \author[UNN,ITMO]{Denis Rodionov}
@@
 
+Дать характеристику фреймворка - язык разработки, используемые технологии, ориентация на какие задачи, и т.п.
+
 \subsection{Global search algorithm}
 \label{sec_GSA}
+
+Алгоримт глобального поиска - ядро фреймворка.
+
+Словесное описание алгоритма.
+
+Формулы, определяющие его работу  (вычисление характеристик, оценка константы Липшица, точка нового испытания).
+
+Описание в виде алгоритма в LaTeX
+
+\begin{algorithm}
+\LinesNumbered
+ \KwIn{this text}
+ \KwOut{how to write algorithm with \LaTeX2e }
+ Initialize $k=1$\\
+ \While{ $\Delta_t \geq \epsilon$ and  $k \leq K_{max}$}{
+  read current\\
+  \eIf{understand}{
+   go to next section\\
+   current section becomes this one
+   }{
+   go back to the beginning of current section
+  }
+ $k = k + 1$\\
+ }
+ \KwRet{$y_{min}$}
+ \caption{Global search}\label{alg_GSA}
+\end{algorithm}
 
@@
 
-А нужно ли тут отдельное описание?
+Здесь описать схему назначения целочисленных значений категориальным параметрам, и схему сведения к задаче  (\ref{problem_is1}).
+
+потом сказать, что задача решается алгоритмом \ref{alg_GSA}.
 
@@
 \label{sec_exp}
+
+\subsection{Experimental setup}
+
+Datasets
+
+Metrics
+
+Algorithms for comparison
+
+\subsection{Experimental results}
+
+First result
+
+Second result
+
+Third result
 

### 2023-07-07 2023/NUMTA/Abstract/Grishagin.tex

--- 2023-04-06 2023/NUMTA/Abstract/Barkalov.tex
@@
 \institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
-\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, vagris@unn.ru}
+\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru, vagris@unn.ru}}
 
@@
 The complexity of these problems increases significantly in the case where criteria are multiextremal.
-In the approaches based on convolution schemes reducing the multicriterial problem to a family of scalar (single criterion) optimization subproblems, these subproblems are multiextremal as well.
+%In the approaches based on convolution schemes reducing the multicriterial problem to a family of scalar (single criterion) optimization subproblems, these subproblems are multiextremal as well.
 This circumstance requires the use of global optimization methods to solve those.
@@
 In the paper a novel algorithm for multicriterial black-box optimization with multiextremal criteria are considered.
-This algorithm applies ideas of complexity reduction when the initial multicriterial problem are reduced to a set of scalar subproblems on the base of maximum convolution, then each scalar multidimensional subproblem is transformed to equivalent one-dimensional problem by means of Peano mapping which provides single-valued correspondence between N-dimensional hyperparallelepiped and the unit interval of the real axis.
-For solving univariate problems a global optimization algorithm with guaranteed convergence to global optimum is used.
+This algorithm applies ideas of complexity reduction when the initial multicriterial problem are reduced to a set of univariate scalar subproblems on the base of Peano mapping and maximum convolution \cite{Gergel2018}.
+%, then each scalar multidimensional subproblem is transformed to equivalent one-dimensional problem by means of Peano mapping which provides single-valued correspondence between $N$-dimensional hyperparallelepiped and the unit interval of the real axis.
+For solving univariate problems a global optimization algorithm with guaranteed convergence to global optimum is used \cite{Sergeyev2013}.
 As a core novel feature, this algorithm includes in its computational scheme a machine learning procedure with combination of accumulating the information of solved subproblems that allows one to significantly accelerate the building the Pareto set.
@@
 \textbf{Acknowledgements.} 
-
 This research was supported by the Russian Science Foundation, project No. 21-11-00204.
@@
 
+\bibitem{Gergel2018}
+Gergel V.P., Kozinov,E.A. (2018) Efficient multicriterial optimization based on intensive reuse of search information. \emph{J. Glob. Optim.},  Vol. 71, pp.~73--90.
+
 \bibitem{Sergeyev2013}
 Sergeyev Y.D., Strongin R.G., Lera D. (2013) \emph{Introduction to global optimization exploiting space-filling curves}. Springer Briefs in Optimization, Springer, New York.
-
-\bibitem{Gergel2018}
-Gergel V.P., Kozinov,E.A. (2018) Efficient multicriterial optimization based on intensive reuse of search information. \emph{J. Glob. Optim.},  Vol. 71, pp.~73--90.
 

### 2023-07-07 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex

@@
 All the competing methods solved the set of 100 MMCO problems several times such that in each case they executed different number of trials and, correspondingly, obtained different values of HV's.
-If we take a method, consider all the pairs $(K, HV(K))$ where $K$ is number of trials executed by the method and $HV(K)$ is the corresponding value of the HV-index, and place all these pairs as points on the plane with coordinate abscissa axis $K$ and ordinate axis $HV$ we build so called \textit{operational characteristic} \cite{Grishagin2016_2} of the method.
+If we take a method, consider all the pairs $(K, HV(K))$ where $K$ is number of trials executed by the method and $HV(K)$ is the corresponding value of the HV-index, and place all these pairs as points on the plane with the coordinate abscissa axis $K$ and the ordinate axis $HV$ we build so called \textit{operational characteristic} \cite{Grishagin2016_2} of the method.
 On the base of operational characteristics it is easy to compare visually the efficiency of different methods.

### 2023-07-07 2023/PAVT_Ufa/PAVT2023.tex

--- 2023-01-29 2023/PAVT_Ufa/PAVT2023.tex
@@
 %
-% RECOMMENDED 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%
-%
+% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{amsmath}
 \usepackage{graphicx}
 \usepackage{marvosym}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{cite}
-
-\usepackage[russian]{babel}
 
@@
 
+%------------------------------------------------------------------------------%
+%-------------------------- Article-specific ----------------------------------%
+%--------------------------    definitions   ----------------------------------%
+%------------------------------------------------------------------------------%
+\newcommand*{\PhUnit}[1]{\phantom{#1}}
+%------------------------------------------------------------------------------%
+
 \begin{document}
 \mainmatter              % start of a contribution
-% Решение обратных задач химической кинетки с помощью асинхронного алгоритма глобального поиска
-% Решение обратных задач химической кинетки с помощью смешанного локально-глобального поискового алгоритма 
-%\title{Solving the Inverse Problems of Chemical Kinetics Using the Asynchronous Global Optimization Algorithm}
-\title{
-Solving ...
-}
-\titlerunning{Solving the Inverse Problems of Chemical Kinetics}  % abbreviated title (for running head)
+%
+\title{Parallel Computing in the Tikhonov Regularization Method for Solving the Inverse Problem of Chemical Kinetics}
+
+%Parallel computing in Tikhonov regularization method for solving the inverse problem of chemical kinetics
+
+%Solving the Regularized Inverse Problem of Chemical Kinetics Using the Parallel Optimization Algorithm
+
+%
+\titlerunning{Parallel Computing in the Tikhonov Regularization Method}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 %
-\author{
-Irek Gubaydullin$^{1,2}$\and
-Leniza Enikeeva$^{2,3}$\orcidID{0000-0003-4219-4870}
-\and
-Konstantin Barkalov$^4$ \letter \orcidID{0000-0001-5273-2471}
-\and
-Marina Usova$^4$\orcidID{XXX} 
-}
+\author{Konstantin Barkalov\inst{1}\letter\orcidID{0000-0001-5273-2471}\and\\ Marina Usova\inst{1}\orcidID{0000-0002-0722-6884} \and\\
+Leniza Enikeeva\inst{2}\letter\orcidID{0000-0003-4219-4870} \and\\
+Dmitry Dubovtsev\inst{3} \and
+Irek Gubaydullin\inst{2,3}}
 
 %
-\authorrunning{I. Gubaydullin et al.} % abbreviated author list (for running head)
+\authorrunning{K. Barkalov et al.} % abbreviated author list (for running head)
 %
 %%%% list of authors for the TOC (use if author list has to be modified)
-%\tocauthor{Konstantin Barkalov and Ilya Lebedev }
+%\tocauthor{Ivar Ekeland and Kim~B.~Bruce, and Roger Temam}
 %
-
-\institute{$^1$Institute of Petrochemistry and Catalysis – Subdivision of the Ufa Federal Research Centre of RAS, Ufa, Russia\\$^2$Ufa State Petroleum Technological University, Ufa, Russia \\$^3$Novosibirsk State University, Novosibirsk, Russia\\$^4$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
-\email{leniza.enikeeva@yandex.ru},
-\email{konstantin.barkalov@itmm.unn.ru},
-\email{ilya.lebedev@itmm.unn.ru}
+\institute{Lobachevsky State University of Nizhny Novgorod,\\
+Nizhny Novgorod, Russian Federation\\
+	\email{konstantin.barkalov@itmm.unn.ru}
+	\and
+	Ufa State Petroleum Technological University,\\
+Ufa, Russian Federation\\
+	\email{leniza.enikeeva@yandex.ru} \\
+        \and
+        Institute of Petrochemistry and Catalysis,\\Subdivision of the Ufa Federal Research Center of\\
+the Russian Academy of Sciences, Ufa, Russian Federation
 }
@@
 \begin{abstract}
+% Ïðîâåðüòå ïåðâîå ïðåäëîæåíèå - ìîé âàðèàíò íèæå
+% The authors investigates the advantages and disadvantages of the products refining processes used in the gasoline production.
+We investigate the advantages and disadvantages of the refining processes employed in gasoline production.
+As a way of increasing the environmental friendliness of motor fuel, we suggest using alkylation to a greater extent during its blending and for improving its quality.
+The work describes the scheme of chemical transformations of the sulfuric acid alkylation process taking into account the target reactions and side effects.
+Based on the chemical nature of the studied reactions, the authors pose the inverse problem of chemical kinetics and consider the regularization method for its solution.
+The global optimization problem corresponding to the regularized inverse problem was solved using a parallel optimization algorithm.
+We provide the results of computational experiments on a supercomputer which show the adequacy of the obtained solution.
+The Tikhonov regularization method is an algorithm intended to find an approximate solution to incorrectly posed operator problems.
+Using this method, the authors solve the task of finding the reaction constants of sulfuric acid alkylation of isoalkanes by alkenes.
 
-The paper considers ...
-%the application of parallel computing technology to the simulation of catalytic chemical reaction, which is widely used in the modern automobile industry to produce gasoline with high octane number.
-As a chemical reaction, the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid is assumed.
-To simulate a chemical process, it is necessary to develop a kinetic model of the process, that is, to determine the kinetic parameters.
-To do this, the inverse problem of chemical kinetics is solved, which predicts the values of kinetic parameters based on laboratory data.
-From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem.
-A parallel information-statistical global search algorithm was used to solve it.
-The use of the parallel algorithm has significantly reduced the search time to find the optimum.
-The found optimal parameters of the model made it possible to adequately simulate the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid.
-
-%обновить ключевые слова
-\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Chemical kinetics $\cdot$ Inverse problems }
+\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Chemical kinetics $\cdot$ Inverse problems $\cdot$ Regularization}
 \end{abstract}
@@
 \section{Introduction}
+%
+To meet the environmental protection requirements, the standards of clean gasoline in the world are currently moving toward low contents of sulfur, olefins, and aromatic substances and high octane numbers.
+This means that the refining industry requires stricter fuel product standards and cleaner production processes.
+Gasoline is a result of the blending of products of several refining processes, such as catalytic reforming, isomerizate, catalytic cracking, and sulfuric acid alkylation.
+The use of each component is limited by certain factors, including sulfur content, saturated vapor pressure, the content of aromatic hydrocarbons, and the octane number of the final product~\cite{alkylation_mod}.
 
-%Часть УГНТУ
-%Currently, there is a tendency to improve the environmental characteristics of automobile fuel while maintaining a high octane number.
-Sulfuric acid alkylation of isobutane with olefins makes it possible to obtain a high-octane component of gasoline with a minimum content of aromatic hydrocarbons.
-The alkylate, which is produced by the alkylation of isobutane with C3-C5 olefins in the presence of strong acid, has the advantages of high octane number, low vapor pressure, and zero content of olefins and aromatics that allow it to be a desirable blending component for high-quality gasoline.
-Alkylates will continue to act as a desirable blending component for high-quality gasoline as the quality of gasoline continues to increase [ссылка на китайскую статью].
-Therefore, it is a significant process of the modern refinery [ссылка на catalyst].
-To optimize the chemical process in industry, it is necessary to first develop its model, which in this case means building a mathematical model of the chemical process and then its kinetic model, that is, numerically calculate the kinetic constants of the reaction.
-%Стыковочный абзац
+The product of catalytic cracking is a catalyst obtained as a by-product of the cracking of vacuum gas oil (fraction 350--500\,\textcelsius).
+It contains aromatic hydrocarbons, olefins, and a small amount of hydrocarbons from the feedstock structure and is a source of sulfur in commercial gasoline.
 
-Since the mathematical model of the chemical reaction is a system of differential equations, it is only possible to find the values of the constants in this system numerically (see, e.g., \cite{RSD}).
-Note that the objective function in such problems is usually multiextremal, i.e.
-it has many local extrema along with a global one.
-%Часть ННГУ
-Numerical methods for solving such multiextremal problems (global optimization methods) differ significantly from local search methods (see, e.g., \cite{Sergeyev2017, PaulaviciusZilinskas2014}).
+An isomerizate of a high-octane component is obtained in the process of catalytic isomerization of light gasoline fractions (i.b.p-62\,\textcelsius).
+The octane number of the isomerizate reaches up to 92 points, depending on the research method.
+However, its excessive addition during blending leads to an increase in the pressure of saturated vapors in commercial gasoline and the formation of steam plugs in the power system during the hot season~\cite{alkylation_mod}.
+
+The source of aromatic contents in gasoline is a reformate obtained during catalytic reforming of heavy gasoline fractions (fraction 85~(105)--180\,\textcelsius).
+The involvement of lighter fractions in reforming feedstocks results in an increase in the proportion of benzene in the reformate.
+
+Alkylated gasoline fully complies with the operational and environmental requirements of modern European, US, and Russian industrial standards for the production of fuels for automotive internal combustion engines and is an ideal and essential component of reformed environmentally friendly gasoline.
+The production rate of alkylate abroad exceeds 70 million tons a year; in the Russian Federation, it reached 2 million tons in 2019.
+Alkylate is obtained as a result of the alkylation of isoalkanes by alkenes in the presence of a catalyst.
+The most commonly used process catalyst in Russia is sulfuric acid.
+The advantages of alkylate include a high octane number (up to 90 points, depending on the research method), low saturated vapor pressure, low content of heteroatomic compounds, and good chemical stability.
+In addition, the sensitivity of alkylate does not exceed 5 points \cite{cao2019}.
+
+There are target and side reactions, which proceed by the carbonium-ion mechanism. The process is carried out in several stages:
+
+1. The first stage is the addition of an acid proton to an olefin to produce a \emph{tert}-butyl carbation:
+\begin{figure}
+\centering
+\includegraphics[height=2.3 cm]{fig1.png}
+\caption{The addition of an acid proton to an olefin to produce a \emph{tert}-butyl carbation}
+\label{fig1}
+\end{figure}
+
+2.
+In the second stage, the formed carbonium ion interacts with paraffin hydrocarbon.
+In this case, the hydrogen anion from the tertiary carbon atom of the isoparaffin hydrocarbon passes into the carbonium ion formed in the first stage of the reaction:
+
+\begin{figure}
+\centering
+\includegraphics[height=3.3 cm]{fig2.png}
+\caption{The interaction of the formed carbonium ion with paraffin hydrocarbon}
+\label{fig2}
+\end{figure}
+
+3. The third stage consists of the addition of a tertiary carbonium ion to the second olefin molecule to form carbonium:
+
+\begin{figure}
+\centering
+\includegraphics[height=3.3 cm]{fig3.png}
+\caption{The addition of a tertiary carbonium ion to the second olefin molecule}
+\label{fig3}
+\end{figure}
+
+4. The fourth stage consists of the skeletal isomerization of the secondary carbonium ion:
+
+\begin{figure}
+\centering
+\includegraphics[height=6.3 cm]{fig4.png}
+\caption{The skeletal isomerization of the secondary carbonium ion}
+\label{fig4}
+\end{figure}
+
+5.
+The fifth stage is the interaction of the formed carbonium ions with an isoparaffin molecule by a tertiary carbon-hydrogen bond with the formation of target products and a carbonium ion from isoparaffin:
+
+\begin{figure}
+\centering
+\includegraphics[height=3.5 cm]{fig5.png}
+\caption{The interaction of the formed carbonium ions with an isoparaffin molecule}
+\label{fig5}
+\end{figure}
+
+For ease of use when compiling the model, we offer below a scheme containing the reactions in the sulfuric acid alkylation process (Fig.\,\ref{fig6}).
+
+\begin{figure}
+\centering
+\includegraphics[height=6.5 cm]{fig6.png}
+\centering
+\caption{The scheme of reactions of the sulfuric acid alkylation process}
+\label{fig6}
+\end{figure}
+
+It should be noted that the temperature range of the process is from 2\,\textcelsius{} to 15\,\textcelsius{}.
+Determining the optimal value for a given composition of feedstocks is one of the study objectives.
+The solution to the problem of constructing a model of the process of sulfuric acid alkylation of isobutane with butylenes is relevant in this regard~\cite{alkylation_mod}.
+
+%Ñòûêîâî÷íûé àáçàö
+Since the mathematical model of the chemical reaction is a system of differential equations, it is only possible to find the values of the constants in this system numerically (see, e.g., \cite{Gubaydullin2021}).
+Note that the objective function in such problems is usually multiextremal, i.e., it has many local extrema apart from the global one.
+
+%×àñòü ÍÍÃÓ
+%The numerical methods for solving such multiextremal problems (global optimization methods) differ significantly from local search methods (see, e.g., \cite{PaulaviciusZilinskas2014,Sergeyev2017}).
 From an algorithmic point of view, they can be divided into two classes: metaheuristic and deterministic.
 Metaheuristic algorithms are mainly based on the simulation of processes occurring in living nature.
-Typical examples of such algorithms are simulated annealing, evolution and genetic algorithms, etc.
-(see, e.g., \cite{Battiti2009,Eiben2015}).
-Due to their relative simplicity, metaheuristic algorithms are more popular with researchers than deterministic ones.
+Typical examples of such algorithms are simulated annealing, evolution and genetic algorithms, and others (see, e.g., \cite{Battiti2009,Eiben2015}).
+Owing to their relative simplicity, metaheuristic algorithms are more popular with researchers than deterministic ones.
 However, the solution to a problem found by a metaheuristic algorithm is, generally speaking, local and may be well away from the global solution \cite{Kvasov2018}.
@@
 Assuming some additional properties of the objective function, it is possible to construct efficient deterministic methods for finding a global solution.
-For example, we can assume that the ratio of the increment of the function to the corresponding increment of the argument cannot exceed some threshold.
-In this case, the functions are called Lipschitz functions, and the problem itself is a Lipschitz global optimization problem.
-This paper continues the series of papers where the parallel methods of Lipschitz optimization proposed in \cite{Strongin2000} are investigated and modified in their application to solving inverse problems of chemical kinetics.
+For example, we can assume that the ratio of the function increment to the corresponding argument increment can not exceed some threshold.
+In this case, the functions are called Lipschitzian, and the problem itself is a Lipschitz optimization problem.
+This paper continues a series of papers where the parallel methods of Lipschitz optimization proposed in \cite{Strongin2000} are investigated and modified in their application to solving inverse problems of chemical kinetics.
 
@@
 Section \ref{Sec_math_mod} describes the mathematical model of the chemical reaction under study.
-The formal statement of the Lipschitz global optimization problem and the general scheme of search algorithms are given in Section \ref{Sec_GO_Problem}.
-In Section \ref{Sec_PGSA}, a scheme of the proposed asynchronous parallel algorithm for solving multiextremal problems is presented.
-The results of numerical solution of the inverse problem of chemical kinetics are discussed in Section \ref{Sec_Exp}.
+The formal statement of the Lipschitz global optimization problem and the general scheme of the search for algorithms are given in Section \ref{Sec_GSA}.
+In that section, we also present a scheme of the proposed asynchronous parallel algorithm for solving multiextremal problems.
+The results of the numerical solution of the inverse problem of chemical kinetics are discussed in Section \ref{Sec_Exp}.
 
-\section{Problem Statement}\label{Sec_math_mod}
-%Содержательная постановка задачи
+\section{The Problem Statement} \label{Sec_math_mod}
 
-\section{Parallel Algorithm for Solving Global Optimization Problems }\label{Sec_GSA}
+To optimize the technological process, it is necessary to know the kinetic laws and mechanisms of chemical reactions.
+To determine them, we should evaluate the kinetic parameters of chemical reactions by solving the inverse problem of chemical kinetics~\cite{regularization}.
+In most cases, the equations of chemical kinetics are systems of ordinary nonlinear differential equations for the concentrations of substances $x_{i}$ (\ref{concentration_changing}) with initial conditions (\ref{init_cond}):
+\begin{gather}
+  \frac{dx_{i}}{dt} = \sum_{j=1} ^J \nu_{ij} \omega_j,\quad i=1,\ldots.I,\ \omega_j=k_j^0 \cdot \exp \biggl(-\frac{E_{j}}{RT}\biggr)\cdot \prod_{i=1} ^M x_{i}^{|a_{ij}|},
+\label{concentration_changing}\\
+t=0,\ x_i(0)=x_i^0.
+\label{init_cond}
+\end{gather}
+Here $k_j^0 \cdot \exp\bigl(-\frac{E_{j}}{RT}\bigr)$ are the so-called rate constants of the reaction stages; $I$ is the number of substances involved in the reaction; $J$ is the number of reaction stages; $\nu_{ij}$ are stoichiometric coefficients; $E_j$ are the activation energies of the reactions (cal/mol); $R$ is the universal gas constant (cal/(mol $\cdot$ K)); $T$ stands for the temperature (K); $a_{ij}$ are the stoichiometric coefficients; $k_j^0$ are pre-exponential multipliers.
 
-\subsection{Global Optimization Problem}\label{Sec_GO_Problem}
+The inverse problem of chemical kinetics is a global optimization problem implying the need to determine the vector of reaction rate constants $(k_1, k_2, \ldots, k_J)$ at which the deviation of the calculated concentrations of the reaction components from the experimental ones is minimal.
+Thus, to determine the rate constants of the reaction, it is necessary to solve the inverse problem of chemical kinetics by repeatedly solving the direct problem, that is, by iterating over the rate constants of the stages (or a set of pre-exponents and activation energies) according to some algorithm.
+To solve the optimization problem, it is necessary to minimize the functional (\ref{functional}), which expresses the deviation of experimental data from the calculated values:
+\begin{equation}
+FF = \sum_{i=1} ^M \sum_{j=1} ^N |x_{ij}^{\mathrm{calc}} - x_{ij}^{\mathrm{exp}}|.
+\label{functional}
+\end{equation}
+Here $x_{ij}^{\mathrm{calc}}$ are the calculated values, $x_{ij}^{\mathrm{exp}}$ are experimental data, $M$ is the number of experimental points, and $N$ is the number of substances involved in the reaction.
+To solve the instability issue in optimization problem (\ref{functional}), consider the minimization problem in the general form
+\begin{equation}
+f(z) \rightarrow \inf,\quad z \in D,
+\label{problem}
+\end{equation}
+where $D$ is some given set and $f\colon D \rightarrow R^1$ is a function (functional) defined on it.
 
-As it was mentioned above, the problem of identifying mathematical model parameter values can be considered a Lipschitz global optimization problem.
-From the formal point of view this problem is a mathematical programming problem of the form
-\begin{gather}
- \varphi^* = \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problemN}\\
- D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\} \label{D},
-\end{gather}
-where $a,b\in R^N$ are the given vectors and $\varphi(y)$ is the objective function which satisfies the Lipschitz condition
-\begin{equation}\label{Lip}
-\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D.
+In this inverse problem of chemical kinetics, a number of kinetic parameters act as a point $z$ (as a rule, a vector of reaction rate constants), while the function $f$ is the deviation of the calculated concentrations of the reaction components from empirical data (functional (\ref{functional})).
+Problem (\ref{problem}) can be attributed to one of two classes of problems, which in the literature are called correctly posed and incorrectly posed minimization tasks, respectively.
+Inverse problems of chemical kinetics are \emph{incorrectly posed problems}  (or, in other words, \emph{ill-posed problems}).
+According to Hadamard, a problem is considered to be \emph{correctly posed} if: 1) its solution exists, 2) the solution is unique, and 3) the solution is stable to variations in the initial data.
+If at least one of the listed requirements does not hold, the task is considered incorrectly posed.
+The inverse problems of chemical kinetics encountered in practice usually have a solution, and condition 1 is, thus, fulfilled.
+Most often, however, conditions 2 and 3 do not hold~\cite{regularization}.
+To solve ill-posed problems approximately, we need special methods to construct a solution to problem (\ref{problem}).
+Such methods exist and are commonly called regularization methods.
+In this paper, we rely on a regularization method suggested by A.\,N.~Tikhonov and called the \emph{stabilization method}.
+Also, we use the classical machine-learning apparatus, i.e., normalization of features and scaling.
+The main task that we will consider is the minimization problem
+\begin{equation}
+f^0(z) \rightarrow \inf,\quad z \in D \subset Z,
+\label{minimization_problem}
+\end{equation}
+in which we assume $D$ to be a convex closed set in a Hilbert space $Z$.
+The function $f^0\colon D \rightarrow R^1$ is considered continuous and convex on $D$.
+Suppose that problem (\ref{minimization_problem}) has a solution, i.e., the set $D^*$ is nonempty.
+Denote by $z^0$ the norm-minimal solution to the problem.
+Suppose that instead of an exact function $f$, we know its approximation $f_{\delta}$, where $\delta \in [0, \delta_0], \delta_0>0$, is a sufficiently small number, and, moreover,
+\begin{equation}
+|f^\delta(z) - f^0(z)| \leq \delta (1+||z||^2),\quad  \forall z \in D.
 \end{equation}
 
-The function $\varphi(y)$ is assumed to be multiextremal and is given as a ``black box'' (i.e.
-as some subroutine with a vector of parameters as its input and the calculated value of the function as its output).
-Moreover, it is assumed that each application of this procedure (hereinafter referred to as search trial) is a time-consuming operation.
+The main construction of the Tikhonov method is the smoothing function or Tikhonov function
+\begin{equation}\label{formula_T}
+T_\alpha^\delta(z) \equiv f^\delta(z) + \alpha \lVert z\rVert^2,\quad z \in D,
+\end{equation}
+where $\alpha > 0$ is the regularization parameter; $\alpha||z||^2$ is called the stabilizing term. Consider the auxiliary minimization problem
+\begin{equation}
+T_\alpha^\delta(z) \rightarrow \inf,\quad z \in D.
+\label{tikhonov_F}
+\end{equation}
+Various numerical methods can be applied to approximate the solution of this problem.
+Suppose that, as a result of a finite number of iterations, we obtain a point $z_\alpha^{\delta,\epsilon}$ such that
+\begin{equation}
+T_\alpha^\delta(z) \equiv \min T_{\alpha}^{\delta}(z) \leq T_{\alpha}^{\delta} (z_{\alpha}^{\delta,\epsilon}) \leq T_{\alpha}^{\delta} + \epsilon,
+\end{equation}
+where $\epsilon > 0$ is a number that characterizes the accuracy of the solution to minimization problem (\ref{tikhonov_F}).
+In this case, minimization problem (\ref{tikhonov_F}) for each fixed pair $\delta, \alpha$ has, as a rule, a much larger ``margin of stability'' than the original problem (\ref{problem}) and, in most cases, is correctly posed.
+
+\section{The Parallel Global Optimization Algorithm}\label{Sec_GSA}
+
+\subsection{The Optimization Problem}\label{Sec_GO_Problem}
+
+As we mentioned above, the task of identifying the model parameter values can be considered a Lipschitz global optimization problem.
+From a formal point of view, this problem is a mathematical programming problem of the form
+\begin{equation} \label{problemN}
+ \varphi^* = \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},
+ \end{equation}
+\begin{equation} \label{D}
+ D=\left\{y\in R^N : a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\},
+ \end{equation}
+where the vectors $a,b\in R^N$ correspond to the lower and upper bounds of the search region, and $\varphi(y)$ is the objective function, which satisfies the Lipschitz condition
+\begin{equation}\label{Lip}
+\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\lVert y_1-y_2\rVert,\quad
+y_1,y_2 \in D.
+\end{equation}
+
+We assume that the function $\varphi(y)$ is multiextremal and is given as a ``black box'' (i.e., as some subroutine with a vector of parameters as its input and the calculated value of the function as its output).
+Moreover, we suppose that each application of this procedure (hereinafter referred to as a \emph{search trial}) is a time-consuming operation.
 This formulation of the problem is fully consistent with the inverse problem of chemical kinetics.
@@
 There are several algorithms that can be applied to solve Lipschitz optimization problems.
-These include, among others, the non-uniform space covering method \cite{Evtushenko2013, Evtushenko2009}, diagonal and simplicial partitions methods \cite{Zilinskas2010, Paulavicius2011}.
-In this paper, we used the global search algorithm proposed by Strongin\cite{Strongin2000}.
+These include, among others, the nonuniform covering method \cite{Evtushenko2009,Evtushenko2013}, and diagonal and simplicial partition methods \cite{Zilinskas2010,Paulavicius2011}.
+In this paper, we rely on the global search algorithm proposed by Strongin \cite{Strongin2000}.
 Under this approach, the original multidimensional problem (\ref{problemN}) is reduced to a one-dimensional optimization problem using Peano-Hilbert curves.
 
-In fact, using a continuous one-to-one mapping (Peano-Hilbert curve) $y(x)$ of the segment $[0,1]$ of the real axis onto the hyperinterval $D$ from (\ref{D}), we can reduce the multidimensional problem (\ref{problemN}) to a one-dimensional problem
+In fact, using a continuous one-to-one mapping (Peano curve) $y(x)$ of the segment $[0,1]$ of the real axis onto the hyperinterval $D$ in (\ref{D}), we can reduce the original problem (\ref{problemN}) to the one-dimensional problem
 \[
@@
 \]
-where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
+where the one-dimensional function $\varphi(y(x))$ satisfies the H{\"o}lder condition
 \[
@@
 with $ H=2 L \sqrt{N+3}$.
-The numerical construction of various approximations of the Peano-Hilbert curve is discussed in \cite{Strongin2000,Sergeyev2013}.
+The numerical construction of various approximations of such mappings is discussed in \cite{Sergeyev2013,Strongin2000}.
 
-Thus, the search trial at some point $x'\in[0,1]$ will involve first the construction of the image $y'=y(x')$, and only then the computation of the function value $ z' = \varphi(y')$.
+Thus, the search trial at some point $x'\in[0,1]$ will involve, in the first place, the construction of the image $y'=y(x')$ and only then the computation of the function value $z' = \varphi(y')$.
 
@@
 
-%Характеристический алгоритм
-The global search algorithm we used belongs to the class of characteristic algorithms, which greatly simplifies the process of its parallelization.
+%Õàðàêòåðèñòè÷åñêèé àëãîðèòì
+The global search algorithm we use belongs to the class of \emph{characteristic algorithms}, and this greatly simplifies the process of its parallelization.
 Recall that a numerical optimization method belongs to the class of characteristic algorithms if the algorithmic scheme of this method can be described as follows.
 
-At the first iteration the trials are performed at the points $x^1 = 0$ and $x^2 = 1$. Then any next $(k+1)$, $k \geq 2$, iteration of global search is executed in the following way.
-
+At the first iteration, the trials are performed at the points $x^1 = 0$ and $x^2 = 1$. Then, each following $(k+1)$-th, $k \geq 2$, iteration of the global search is executed as described below.
 \begin{enumerate}
 	\item
-Renumber points of the previous trials by increasing their $x$-coordinates (the new order is marked with subscript)
+Renumber the points of the previous trials by increasing their $x$-coordinates (the new order is determined by the subscripts):
 \[
-0=x_0<x_2<...<x_{k}=1.
+0=x_0<x_2<\ldots<x_{k}=1.
 \]
@@
 	\item
-Calculate for every interval $(x_{i-1},x_i), 1\leq i\leq k$, a value $R(i)$ called \textit{the interval characteristic}.
-In the general case, $R(i)$ can depend on the points $x^i$ and the trial results $z^i=f(x^i), 1 \leq i \leq k$.
+For every interval $(x_{i-1},x_i), 1\leq i\leq k$, we calculate a value $R(i)$ called the \textit{interval characteristic}.
+In the general case, $R(i)$ can depend on the points $x^i$ and the trial results $z^i=f(x^i)$, $1 \leq i \leq k$.
 
 	\item
-Find the interval $(x_{t-1},x_t)$ which has the largest characteristic $R(t)$, i.e.
+Find the interval $(x_{t-1},x_t)$ with the largest characteristic $R(t)$, i.e.,
 \[
-R(t) = \max \left\{ R(i): \; 1\leq i\leq k \right\}.
+R(t) = \max \left\{ R(i) : 1\leq i\leq k \right\}.
 \]
@@
 \]
-where $\epsilon>0$ is a given accuracy. If the stop condition is satisfied, then the global search has to be terminated; otherwise the calculations should continue.
+where $\epsilon>0$ is a given accuracy. If the stop condition is satisfied, then the global search should be terminated; otherwise, the calculations continue.
 
 	\item
-Select a point $x^{k+1}$ of the current iteration within the interval $(x_{t-1},x_t)$ in accordance with some rule $S(t)$, i.e.
+Select a point $x^{k+1}$ of the current iteration within the interval $(x_{t-1},x_t)$ in accordance with some rule $S(t)$, i.e.,
 \[
@@
 	\item
-Calculate the function value $z^{k+1} = f(x^{k+1})$
+Calculate the function value $z^{k+1} = f(x^{k+1})$.
 
 	\item
-Evaluate a global minimum estimate
+Evaluate a global minimum estimate.
 
@@
 
-As a possible interpretation of this scheme, we can consider an interval characteristic $R(i)$ as a measure of the global minimum being within the interval
+As a possible interpretation of this scheme, we can consider the interval characteristic $R(i)$ as a measure of the global minimum being within the interval
 $(x_{i-1},x_i)$.
-To construct a concrete form of the interval characteristic, we can use a lower envelope (or minorant) of the function to be minimized or a mathematical expectation of function values, etc. 
-Most well-known global optimization methods can be formulated in accordance with this characteristical scheme, e.g.,
+To construct a concrete form of the interval characteristic, we can use a lower envelope (or minorant) of the function to be minimized or a mathematical expectation of the function values, etc.
+Most well-known global optimization methods can be formulated in accordance with this characteristic scheme, e.g.,
 \begin{itemize}
-\item a uniform grid method with a successively reduced step;
-\item a random search (Monte-Carlo) algorithm;
-\item the Piyavskii method;
-\item one-step bayesian methods proposed by Kushner and Zilinskas;
-\item information algorithms proposed by Strongin.
+\item uniform grid methods with a successively reduced step;
+\item random search (Monte Carlo) algorithms;
+\item the Piyavskii algorithm \cite{Piyavskii1972};
+\item one-step Bayesian algorithms proposed by Kushner and \v{Z}ilinskas \cite{Zilinskas1989};
+\item information algorithms proposed by Strongin \cite{Strongin2000}.
 \end{itemize}
-All these methods are based on different mathematical models but are presented in the general characteristical scheme.
+All these methods are based on different mathematical models but are presented in a general characteristic scheme.
 
-It should be also noted that the length of the interval with the largest characteristic is examined at the stop condition.
+It should also be noted that the length of the interval with the largest characteristic is examined at the stop condition.
 It is possible if the optimization method converges to the global minimum.
+The details of these methods are given in the cited sources.
+%Çäåñü æå îòìåòèì îñíîâíûå ìîìåíòû
+Here we highlight the main points.
 
-Details of these methods are given in the sources cited. %Здесь же отметим основные моменты
+For the Piyavskii method, the interval characteristic $R(i)$ is an estimate (with the inverse sign) of the minimum value of the objective function $f(x)$ in the interval $(x_{i-1},x_i)$.
+As a result, the point of a new trial is taken within the interval containing the estimate of the minimum value of $f(x)$ over the search domain.
 
-For the Piyavskii method the interval characteristic $R(i)$ is an estimate (with the inverse sign) of the least value of the objective function $f(x)$ within the
-interval $(x_{i-1},x_i)$. As a result, the point of a new trial is taken within the interval containing the estimate of the least value of $f(x)$ over the search domain.
-
-The Kushner technique and the Zilinskas method have been constructed in the framework of the approach when the objective function is considered as
-a sample of some Wiener process. For the Kushner technique the point $x^{k+1}$ of the current iteration is the most probable point at which the function value
+The Kushner technique and the \v{Z}ilinskas method are constructed in the framework of the approach when the objective function is regarded as
+a sample of some Wiener process. For the Kushner technique, the point $x^{k+1}$ of the current iteration is the most probable point at which the function value
 $f(x^{k+1})$ is not greater than the value
@@
 \]
-where $z_k^+$ and $z_k^*$ are estimates of maximum and minimum function values respectively.
-For the Zilinskas method $x^{k+1}$ is the point where the maximum average improvement of the current estimate of the global extremum is expected.
+where $z_k^+$ and $z_k^*$ are estimates of the function maximum and minimum values, respectively.
+For the \v{Z}ilinskas method, $x^{k+1}$ is the point where the maximum average improvement of the current estimate of the global extremum is expected.
 
-The Strongin global search algorithm has been constructed in the framework of the information approach to global optimization (see \cite{Strongin2000})).
-This method has an adaptive scheme to evaluate a numerical estimate of the unknown Lipschitz constant.
+The Strongin global search algorithm is constructed in the framework of the information approach to global optimization (see \cite{Strongin2000}).
+This method has an adaptive scheme to evaluate the numerical estimate of the unknown Lipschitz constant.
 
-\subsection{Parallel Asynchronous Global Search Algorithm}\label{Sec_PGSA}
+\subsection{Parallel Algorithm with Asynchronous Trials}\label{Sec_PGSA}
 
-As the global search algorithm (GSA) belongs to the class of characteristic algorithms, this suggests a possible way to parallelize it. 
-As previously mentioned, the characteristic of the interval $R(i)$ can be regarded as some measure of finding the global minimum point in this interval.
-Then, instead of a single best interval, several intervals with the highest characteristics can be chosen at once and successive trials can be carried out in these intervals in parallel.
-Moreover, the scheme of characteristic algorithms also allows asynchronous parallelization, which minimizes downtime of processors when trial complexity depends on a particular point in the search domain.
+The global search algorithm (GSA) belongs to the class of characteristic algorithms; this fact suggests a possible way to parallelize it.
+As previously mentioned, the interval characteristic $R(i)$ can be regarded as some measure of finding the global minimum point in the given interval.
+Then, instead of a single best interval, several intervals with the highest characteristics can be chosen at once, and successive trials can be carried out in these intervals in parallel.
+Moreover, the scheme of characteristic algorithms also allows for asynchronous parallelization, which minimizes the downtime of processors when the trial complexity depends on a particular point in the search domain.
 
-Now, let us examine the asynchronous global search algorithm in more detail.
-It implements a paralleling scheme of the ``master/worker'' type.
-The master process accumulates search information, evaluates on its basis the Lipschitz constant for the objective function, determines new trial points and distributes them to the worker processes.
-Worker processes receive the points from the master, perform new trials in these points and send the results to the master.
+Let us now examine the parallel algorithm with asynchronous trials in more detail.
+It implements a parallel scheme of the ``master/worker'' type.
+The master process accumulates search information, evaluates on its basis the Lipschitz constant for the target function, determines new trial points, and sends them to the worker processes.
+Worker processes receive points from the master, carry out new trials at these points, and send the results to the master.
 
-When describing the parallel algorithm, we assume that we have $p+1$ computational processes: one master and $p$ worker processes.
+When describing the parallel algorithm, we assume that there are $p+1$ computational processes: one master and $p$ worker processes.
 
-At the beginning of the search, the master process (let it be process No.
-0) initiates $p$ parallel  trials at $p$ different points of the search domain, two of which are boundary points and the rest are internal points, i.e.
-at the points $\{y(x^1), y(x^2), ...,y(x^p)\}$, where $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.
+At the beginning of the search, the master process (let it be process number 0) initiates $p$ parallel trials at $p$ different points of the search domain, two of which are boundary points and the rest are internal points, i.e., at the points $\{y(x^1), y(x^2), \ldots,y(x^p)\}$, where $x^1 = 0$, $x^p = 1$, $x^i\in(0,1)$, $i=2,\ldots, p-1$.
 
-Suppose now that $k$ trials have been performed (in particular, $k$ can be 0), and the worker processes perform trials at the points $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.
+Suppose now that $k$ trials have been performed (in particular, $k$ can be 0), and the worker processes perform trials at points $\{y(x^{k+1}), y(x^{k+2}), \ldots,y(x^{k+p})\}$.
 
-If a worker process completes the trial at some point (let it be the point $y(x^{k+1})$ corresponding to process No.
-1) then it sends to the master process the results of the trial.
-Note that in this case we will have a set of preimages of the trial points
+If a worker process completes the trial at some point (let it be the point $y(x^{k+1})$, corresponding to process number 1), then it sends the results of the trial to the master process.
+Note that, in this case, we have a set of inverse images of the trial points
 \[
-I_k = \left\{ x^{k+1},x^{k+2},...,x^{k+p} \right\},
+S_k = \left\{ x^{k+1},x^{k+2},\ldots,x^{k+p} \right\},
 \]
-where the trials have already started but have not yet been completed.
+where the trials have already started but are not completed yet.
 
-Having received the trial results at the point $y(x^{k+1})$ from the worker process, the master in turn selects a new trial point $x^{k+p+1}$ for it according to the rules that correspond to the scheme of the characteristic algorithm.
-
+After receiving the trial results at the point $y(x^{k+1})$ from the worker process, the master selects a new trial point $x^{k+p+1}$ for it, according to the rules corresponding to the scheme of the characteristic algorithm.
 \begin{enumerate}
 	\item
-Renumber the set of preimages of the trial points 
+Renumber the set of inverse images of the trial points
 \[
-X_k = \left\{x^1, x^2,...,x^{k+p} \right\},
+X_k = \left\{x^1, x^2,\ldots,x^{k+p} \right\},
 \]
-containing all preimages at which the trials have either been carried out or are being carried out in ascending order (marked with subscript) , i.e.
+which contains all inverse images at which the trials have either been carried out or are being carried out, arranged in an ascending order (determined by the subscripts), i.e.,
 \[
-0=x_1<x_2<...<x_{x+p}=1.
+0=x_1<x_2<\ldots<x_{x+p}=1.
 \]
+
 	\item
@@
 \[
-M_1=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin I_k, x_i \notin I_k, 2\leq i\leq k+p \right\},
+\begin{aligned}
+M_1&=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin S_k, x_i \notin S_k, 2\leq i\leq k+p \right\},\\
+M_2&=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in S_k, 2\leq i < k+p \right\},\\
+M&=\max\{M_1,M_2\},
+\end{aligned}
 \]
-\[
-M_2=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in I_k, 2\leq i < k+p \right\},
-\]
-\[
-M=\max\{M_1,M_2\},
-\]
-where $z_i=\varphi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$.
-The values of $z_i$ at the points $x_i \in I_k$ are undefined, as the trials at the points $x_i \in I_k$ have not yet been completed.
-If the value of $M$ equals 0, assume $M=1$.
+where $z_i=\varphi(y(x_i))$ for all $x_i \notin S_k$.
+%The values of $z_i$ at the points $x_i \in S_k$ are undefined, as the trials at these points have not yet been completed.
+If $M$ equals 0, then assume $M=1$.
 
 	\item
-Assign to each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, the number $R(i)$, which is called the interval characteristic and is calculated using the formula
+Assign to each interval $(x_{i-1},x_i)$, $x_{i-1} \notin S_k$, $x_i \notin S_k$, the characteristic $R(i)$, calculated by the formula
 \begin{equation} \label{R}
@@
 \end{equation}
-where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.
+where $r>1$ is a method parameter and $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$.
 
 	\item
-Select the interval $[x_{t-1},x_t]$ to which the maximum characteristic corresponds, i.e.
+Find the interval $[x_{t-1},x_t]$ to which the maximum characteristic corresponds, i.e.,
 \[
-R(t) = \max \left\{ R(i): \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p \right\}.
+R(t) = \max \left\{ R(i) : x_{i-1} \notin S_k, x_i \notin S_k, 2\leq i\leq k+p \right\}.
 \]
+
 	\item
-Determine the preimage $x^{k+p+1} \in (x_{t-1},x_t)$ of the new trial point according to the formula
+Determine the inverse image $x^{k+p+1} \in (x_{t-1},x_t)$ of the new trial point by the formula
 \[
-x^{k+p+1} = \frac{x_{t}+x_{t-1}}{2} - \mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_{t}-z_{t-1}\right|}{M}\right]^N.
+x^{k+p+1} = \frac{x_{t}+x_{t-1}}{2} - \operatorname{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_{t}-z_{t-1}\right|}{M}\right]^N.
 \]
@@
 
-Immediately after calculating the next trial point $y^{k+p+1} = y(x^{k+p+1})$, the master process adds it to the set $I_k$ and forwards it to the worker process which initiates a trial at this point.
+Immediately after calculating the next trial point $y^{k+p+1} = y(x^{k+p+1})$, the master process adds it to the set $S_k$ and forwards it to the worker process, which initiates a trial at this point.
 
-The master process completes the algorithm when one of two conditions is met: $\Delta_{t}<\epsilon$, $k+p>K_{max}$.
-). The first one corresponds to stopping the algorithm by accuracy, the second one, by the number of trials. The real number $0<\epsilon<1$ and the integer $K_{max}>0$ are the algorithm parameters.
+The master process completes the algorithm when one of two conditions holds: $\Delta_{t}<\epsilon$ or $k+p>K_{\mathrm{max}}$.
+The first one corresponds to stopping the algorithm by accuracy, and the second one, by the number of trials.
+The real number $0<\epsilon<1$ and the integer $K_{\mathrm{max}}>0$ are algorithm parameters.
 
-\section{Numerical Experiments}\label{Sec_Exp}
+\section{Numerical Experiments} \label{Sec_Exp}
 
-In preliminary experiments, we estimated the approximate time to calculate one objective function using the sequential algorithm.
-The computer architecture used for the experiments was Intel(R) Core(TM) i7-10750H CPU @ 2.60GHz 2.59 GHz.
-The average calculation time of the objective function was about $0.2$ sec, which characterizes the problem as being computationally complex.
-Solving the problem with a sequential algorithm would require at least 1 million trials and thus at least 54 hours (as an estimate).
-Therefore only a parallel algorithm with a stopping criterion based on the number of trials $K_{max}=10^6$ was used for the calculations.
+To determine the constants of the process reactions, it is necessary to solve a system of differential equations that describes the change in the concentrations of the reaction components over time \cite{cao2019}.
 
-Numerical experiments were performed using the parallel asynchronous algorithm outlined in Section \ref{Sec_PGSA}, running $160$ processes: $159$ processes calculated the values of the objective function (worker processes) while $1$ master process controlled the algorithm.
-The Lobachevsky  supercomputer (University of Nizhni Novgorod)  (CentOS 7.2, SLURM, two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM on the node) was used for the experiments.
-The asynchronous global optimization algorithm was implemented in C++ (GCC 9.5.0 and Intel MPI were used); the computation of objective function values was implemented in Python 3.9.
+In preliminary experiments, we estimated the approximate time to calculate one objective function using the corresponding sequential algorithm.
+The computer architecture used for the experiments was based on Intel\textsuperscript{\tiny\textregistered} Core\textsuperscript{\tiny\texttrademark} i7-10750H processors at 2.60~GHz.
+The average computation time of the objective function was about $0.2$ sec, which characterizes the problem as being computationally complex.
+Solving the problem with the sequential algorithm would require no less than 1 million trials, that is, at least 54 hours (as an estimate).
+Therefore, only the parallel algorithm with a stopping criterion based on the number of trials $K_{\mathrm{max}}=10^6$ was used for the calculations.
 
-%Ускорение составило примерно $230$ раз.
-The short time for solving the problem when using the parallel algorithm (approximately 14 minutes for 1 problem) allowed a thorough investigation of the dependence of the solution on the regularisation parameter $\alpha$.
-The table \ref{table_1} shows the minimum values of the objective function found for the corresponding value of the regularization parameter, the time for solving the problem (in seconds) and the time speedup in comparison to the estimated problem solution time when using the sequential method.
-The method parameter $r=4.0$ from (\ref{R}) and the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$ were used in the runs.
-After a given number of iterations or after reaching the specified accuracy with the global search method, the solution was refined by the Hooke-Jeeves local method \cite{HookJeeves} with an accuracy of $\epsilon = 10^{-4}\left\|b-a\right\|$.
-
-A detailed study of the region of the regularization parameter values in the neighborhood of $\alpha$ equal to zero (from $0.1$ to $0.01$ in steps of $0.01$ and from $0.01$ to $0.001$ in steps of $0.001$) revealed that the value of the minimum decreases with decreasing $\alpha$.
-The best solution of $0.54075$ was obtained with the regularisation parameter $\alpha = 0.001$.
+The numerical experiments were performed with the parallel asynchronous algorithm outlined in Section \ref{Sec_PGSA}, running 160 processes: 159 processes calculated the values of the objective function (worker processes), and one master process controlled the algorithm.
+The experiments were conducted on the Lobachevsky supercomputer, installed at the University of Nizhni Novgorod.
+The asynchronous global optimization algorithm was implemented in C++ (with GCC 9.5.0 and Intel MPI).
+The computation of the objective function values was implemented in Python 3.9.
 
@@
 \begin{center}
-\begin{tabular}{ccc}
+\begin{tabular}{c@{\quad}cc}
 \hline\noalign{\smallskip}
@@
 \hline\noalign{\smallskip}
-0.1		&	0.810454	&	806.192	\\
-0.09	&	0.837868	&	828.649 \\
-0.08	&	0.808238	&	811.852	\\
-0.07	&	0.759134	&	803.626 \\
-0.06	&	0.723404	&	810.9		\\
-0.05	&	0.729415	&	804.674	\\
-0.04	&	0.736035	&	816.387	\\
-0.03	&	0.68152		&	809.185	\\
-0.02	&	0.669092	&	155.837	\\
-0.01	&	0.632944	&	825.238	\\
-0.009	&	0.63001		&	814.48	\\
-0.008	&	0.604817	&	830.448	\\
-0.007	&	0.619546	&	807.335	\\
-0.006	&	0.612614	&	834.703	\\
-0.005	&	0.615638	&	815.45	\\
-0.004	&	0.579725	&	821.421	\\
-0.003	&	0.583529	&	831.454	\\
-0.002	&	0.547907	&	850.017	\\
-0.001	&	0.54075		&	853.898	\\
+0.0\PhUnit{00} &	0.573920 & 838.9	\\
+0.001	&	0.540750	&	853.9	\\
+0.002	&	0.547907	&	850.0	\\
+0.003	&	0.583529	&	831.5	\\
+0.004	&	0.579725	&	821.4	\\
+0.005	&	0.615638	&	815.4	\\
+0.006	&	0.612614	&	834.7	\\
+0.007	&	0.619546	&	807.3	\\
+0.008	&	0.604817	&	830.4	\\
+0.009	&	0.630010	&	814.4	\\
+0.01\PhUnit{0}	&	0.632944	&	825.2	\\
+0.02\PhUnit{0}	&	0.669092	&	155.8	\\
+0.03\PhUnit{0}	&	0.681520	&	809.2	\\
+0.04\PhUnit{0}	&	0.736035	&	816.4	\\
+0.05\PhUnit{0}	&	0.729415	&	804.7	\\
+0.07\PhUnit{0}	&	0.759134	&	803.6 \\
+0.06\PhUnit{0}	&	0.723404	&	810.9	\\
+0.08\PhUnit{0}	&	0.808238	&	811.8	\\
+0.09\PhUnit{0}	&	0.837868	&	828.6 \\
+0.1\PhUnit{00}	&	0.810454	&	806.2	\\
 \noalign{\smallskip}\hline
@@
 
-\section{Conclusions and Future Work}
+%Óñêîðåíèå ñîñòàâèëî ïðèìåðíî $230$ ðàç.
+The short time obtained when solving the problem by the parallel algorithm (approximately 14 minutes per problem) made it possible to investigate thoroughly the dependence of the solution on the regularization parameter $\alpha$.
+Table \ref{table_1} shows the minimum values of the objective function found for the corresponding value of the regularization parameter and the time for solving the problem (in seconds).
+%and the time speedup in comparison to the estimated problem solution time when using the sequential method.
+The algorithm was run with the method parameter $r=4.0$ from (\ref{R}) and the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$.
+After a given number of iterations or after reaching the specified accuracy with the global search method, the solution was refined by the Hooke--Jeeves local method \cite{HookJeeves} with the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$.
 
-\medskip
+A detailed study of the region of the regularization parameter values in a neighborhood of $\alpha=0$ (from $0.1$ to $0.01$ with steps of $0.01$ and from $0.01$ to $0.001$ with steps of $0.001$) showed that the value of the minimum decreases as $\alpha$ decreases.
+The best solution, equal to $0.54075$, was obtained with the regularization parameter $\alpha = 0.001$.
 
-\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204 and by RFBR, project No.\,19-37-60014.
+In the course of the research, we calculated the functional $f(z)$ without taking into account the addition of the regularization term (Fig.\,\ref{fig7}).
+Afterward, we carried out a similar calculation taking into account the regularization term.
+The results for $\alpha=0.002$ were $0.527$ without regularization and $0.547$ with regularization.
+
+\begin{figure}
+\centering
+\includegraphics[height=6.0 cm]{fig7.png}
+\caption{The value of the functional $f(z)$ without taking into account the addition of the regularization term}
+\label{fig7}
+\end{figure}
+
+The speedup of the parallel global optimization algorithm was estimated for the problem to which the best solution was found, i.e., the problem with regularization parameter $\alpha = 0.001$.
+Since solving the problem in serial mode would require about two days, the speedup was evaluated with respect to a parallel launch involving a smaller number of processes.
+Table \ref{table_2} shows the running time and speedup of the parallel algorithm using 80 and 160 processes with respect to the running time on 40 processes.
+
+\begin{table}
+\caption{The speedup of the parallel global optimization algorithm}
+\label{table_2}
+\begin{center}
+\begin{tabular}{ccc}
+\hline\noalign{\smallskip}
+ $p$     & Time (sec.)  & Speedup \\
+\hline\noalign{\smallskip}
+\PhUnit{0}40	&	2778.7		&	--	\\
+\PhUnit{0}80	&	1495.5		&	1.9	\\
+160	&	\PhUnit{0}853.9	    &	3.3	\\
+\noalign{\smallskip}\hline
+\end{tabular}\end{center}\end{table}
+
+\section{Conclusions}
+
+We solved in this paper an incorrectly posed problem using the Tikhonov regularization method and found the reaction constants of the process of sulfuric acid alkylation of isoalkanes by alkenes.
+The optimization method allowed us to find an accurate description of the experimental data.
+The authors plan to use the Tikhonov regularization method in forthcoming research to build kinetic models of other chemical processes.
+
+The parallel global search algorithm showed good results.
+The solution to the problem was obtained in approximately 15 minutes on the Lobachevsky supercomputer; the estimated time for solving the same problem with the corresponding sequential algorithm was more than 50 hours.
+The experiments were conducted using 160 processes on the supercomputer nodes.
+However, an increase in the dimension of the considered global optimization problem (in the case of complex inverse problems of chemical kinetics, the number of parameters can be hundreds) leads to a decrease in search quality.
+
+A possible topic for further research is the development of methods for analyzing the accumulated search information to identify groups of parameters that have little effect on the objective function.
+To find the optimal values of such parameters, it is sufficient to employ local optimization methods.
+The analysis can rely, for example, on machine-learning methods.
+In this case, the cost of solving the whole problem decreases, and it becomes feasible to obtain the global solution with good accuracy in an acceptable time.
+
+\subsubsection*{Acknowledgments.}
+%This study was supported by the Ministry of Science and Higher Education of the Russian Federation (project no.
+FSWR-2023-0034), and by the Research and Education Mathematical Center (project no.
+075-02-2022-883).
+This research was partially funded by the state task of the Institute of Petrochemistry and Catalysis of the Russian Academy of Sciences (theme No.
+FMRS-2022-0078).
+
+%This research was partially funded by the Ministry of Science and Higher Education of the Russian Federation (project \textnumero~FSWR-2023-0034), the Research and Education Mathematical Center (project \textnumero~075-02-2022-883; development of the parallel global optimization algorithm), and the Ministry of Science and Higher Education of the Russian Federation (project \textnumero~FMRS-2022-0078; investigation of the chemical kinetics problem).
+
+This research was partially funded by the projects \textnumero~FSWR-2023-0034, \textnumero~075-02-2022-883 (development of the parallel global optimization algorithm), and by the project \textnumero~FMRS-2022-0078 (investigation of the chemical kinetics problem).
 

### 2023-07-09 2023/ITMO paper/elsarticle-template-num.tex
test algorithm

--- 2023-06-14 2023/ITMO paper/elsarticle-template-num.tex
@@
 
+\begin{algorithm}
+ \LinesNumbered
+ \KwIn{this text}
+ \KwOut{how to write algorithm with \LaTeX2e }
+ Initialize $k=1$\\
+ \While{ $\Delta_t \geq \epsilon$ and  $k \leq K_{max}$}{
+  Calculate the estimate of the H\"older constant $m$ from (\ref{eq:19}) using $A_k$\\
+	\For{$i=1$ \KwTo $k$}{
+	g\\
+	}
+
+  \eIf{understand}{
+   go to next section\\
+   current section becomes this one
+   }{
+   go back to the beginning of current section
+  }
+ $k = k + 1$\\
+ }
+ \KwRet{$y_{min}$}
+ \caption{Global search}
+\end{algorithm}
+
 \subsection{GSA for the problems with categorical parameters}

### 2023-07-27 2023/vestnik.susu/example.tex
Статья в вестник Южно-Уральского государственного университета

--- 2023-02-01 2023/PAVT/ParallelDecisionTree.tex
@@
-%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% sample root file for your "contribution" to a proceedings volume
-%
-% Use this file as a template for your own input.
-%
-%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass{cmi}
+\usepackage{upgreek}
 
-\documentclass{svproc}
-\usepackage{marvosym}
-%
-% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\usepackage{graphicx}
+% Данные поля заполняются редакцией журнала
+%\makeatletter
+%\newcommand{\@volume}{}
+%\newcommand{\@issue}{}
+%\newcommand{\@num}{}
+%\newcommand{\@pages}{}
 
-% to typeset URLs, URIs, and DOIs
-\usepackage{url}
-\usepackage{hyperref}
-\usepackage[english]{babel} 
-\usepackage{amsmath} 
-\def\UrlFont{\rmfamily}
-\providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}
-
-\def\orcidID#1{\unskip$^{[#1]}$}
-\def\letter{$^{\textrm{(\Letter)}}$}
+% Подключение bib файлов
+\addbblresources{biblio_ru.bib}{biblio_en.bib}
 
 \begin{document}
-\mainmatter              % start of a contribution
-%
-\title{On using the decision trees to identify the local extrema in parallel global optimization algorithms}
-%
-\titlerunning{On using the decision trees to identify the local extrema }  % abbreviated title (for running head)
-%                                     also used for the TOC unless
-%                                     \toctitle is used
-%
-\author{Konstantin Barkalov\inst{1}\letter\orcidID{0000-0001-5273-2471} \and Ilya Lebedev\inst{1}\letter\orcidID{0000-0002-8736-0652} \and Dmitry Silenko\inst{1}\letter\orcidID{0000-0002-2578-9699}}
-%
-\authorrunning{Konstantin Barkalov et al.} % abbreviated author list (for running head)
-%
-%%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{K.A.~Barkalov and I.G.~Lebedev, and D.I.~Silenko}
-%
-\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
-\email{ \{konstantin.barkalov,ilya.lebedev,silenko\}@itmm.unn.ru}
+
+% ЗАПОЛНЯЕТСЯ АВТОРАМИ: Релевантные коды УДК статьи
+% (см. http://teacode.com/online/udc/)
+\classify{УДК  519.853.4}
+
+\renewcommand*{\thefootnote}{\fnsymbol{footnote}} %to make a number in end of title delete this line
+
+\title{Об использовании деревьев решений для выявления областей притяжения локальных минимумов в параллельном алгоритме глобальной оптимизации%
+	\footnote{Статья рекомендована к публикации программным комитетом Международной научной конференции <<Параллельные вычислительные технологии (ПаВТ) 2023>>}%
 }
 
-\maketitle              % typeset the title of the contribution
+\authors{К.А.~Баркалов$^{1}$, И.Г.~Лебедев$^{1}$, Д.И.~Силенко$^{1}$}
 
-\begin{abstract}
+\organizations{$^{1}$Нижегородский государственный университет им. Н.И. Лобачевского\\
+	(603022, Н.Новгород, проспект Гагарина, ~23)}
+
+\newcommand{\emailsinsert}{ \href{mailto:barkalov@vmk.unn.ru}{barkalov@vmk.unn.ru}, \href{mailto:ilya.lebedev@itmm.unn.ru}{ilya.lebedev@itmm.unn.ru}, \href{mailto:silenko@itmm.unn.ru}{silenko@itmm.unn.ru} }
+
+% ЗАПОЛНЯЕТСЯ АВТОРАМИ: Дата поступления статьи в редакцию 
+\newcommand{\rdate}{28.07.2023}
+\emails{\emailsinsert}
+\redactiondate{\rdate}
+
+\renewcommand*{\thefootnote}{\arabic{footnote}}
+\setcounter{footnote}{0}
+
+% Оформление аннотации
+\begin{abstract}%
+В работе рассматривается решение многомерных задач глобальной оптимизации с использованием деревьев решений для выявления областей притяжения локальных минимумов.
+Целевая функция представлена как «черный ящик», для которой известно, лишь что она удовлетворяет условию Липшица с неизвестной константой.
+Для решения этой задачи применятся алгоритм глобального поиска.
+В работе предлагается метод определения окрестности локальных минимумов целевой функции на основе анализа накопленной поисковой информации.
+Проведение такого анализа с использованием методов машинного обучения позволяет принять решение о запуске локального метода, что может ускорить сходимость алгоритма.
+Данный подход был подтвержден результатами численных экспериментов, демонстрирующих ускорение при решении набора тестовых задач.
+
+\keywords{Глобальная оптимизация, Многоэкстремальные функции, Параллельные вычисления, Машинное обучение, Дерево решений}
+\end{abstract}
+
+%Образец цитирования статьи
+\forcitation{Баркалов К.А., Лебедев  И.Г., Силенко  Д.И. Об использовании деревьев решений для выявления областей притяжения локальных минимумов в параллельном алгоритме глобальной оптимизации}
+
+%Верхние колонтитулы
+\markboth{К.А.~Баркалов, И.Г.~Лебедев, Д.И.~Силенко}{Об использовании деревьев решений в параллельном алгоритме глобальной оптимизации \ldots}
+
+\vspace{1em}
+{\it Работа выполнена при поддержке Российского научного фонда, проект № 21-11-00204 .}
+
+\insertbiblio
+
+% Сведения об авторах (Ф.И.О.(полностью), степень, звание, кафедра (отдел), вуз, e-mail)
+Первый Александр Васильевич, д.ф.-м.н., профессор, кафедра системного программирования, Южно-Уральский государственный университет (национальный исследовательский университет) (Челябинск, Российская Федерация)
+
+Второй Владимир Григорьевич, к.т.н., доцент, кафедра суперкомпьютеров и квантовой информатики, Московский государственный университет имени М.В.~Ломоносова (Москва, Российская Федерация)
+
+%\newpage
+
+{\phantom{eee} \hrule \vskip 4 pt}
+\classify{} % Код MSC не нужен
+%\classify{MSC 00000} % Коды MSC см. http://www.ams.org/msc/
+
+\title{On using the decision trees to identify the local extrema in parallel global optimization algorithms}
+
+\authorsen{Konstantin Barkalov$^{1}$, Ilya Lebedev$^{1}$,  Dmitry Silenko$^{1}$}
+
+\organizations{$^{1}$Lobachevsky State University of Nizhny Novgorod\\
+	(23 Prospekt Gagarina, Nizhny Novgorod, 603022 Russia)}
+
+\emails{\emailsinsert}
+\redactiondateen{\rdate}
+
+\begin{abstract}%
 In the present work, the solving of the multidimensional global optimization problems using decision tree to reveal the attractor regions of the local minima is considered.
@@
 This suggestion was confirmed by the results of numerical experiments demonstrating the speedup  when solving a series of test problems.
-% We would like to encourage you to list your keywords within
-% the abstract section using the \keywords{...} command.
-\keywords{Global optimization $\cdot$ Multiextremal functions $\cdot$ Parallel computing $\cdot$ Machine learning $\cdot$ Decision tree}
+
+\keywordsen{Global optimization, Multiextremal functions, Parallel computing, Machine learning, Decision tree}
 \end{abstract}
-%
 
+\forcitationen{Pervyi~A.B., Vtoroi~V.G. The \LaTeXe{} Template for Paper Submission for the ``Computational Mathematics and Software Engineering'' Series}
+
+% Заголовки разделов формируются при помощи команд  \section{}, \subsection{}, \subsubsection{}
 \section{Introduction}
@@
 Such problems often arise in the cases, when it is necessary to select the values  of parameters of the mathematical model being investigated, in which the simulation results best fit the experimental data.
-For solving the problems of this class, many algorithms are known: from the metaheuristic algorithms, based on the  idea of random search \cite{Ferreiro2013,Garcia2014,Langdon2011}, to the deterministic algorithms guarantying convergence to the global minimum \cite{Evtushenko2009,He2008,Paulavicius2011}.
+For solving the problems of this class, many algorithms are known: from the metaheuristic algorithms, based on the  idea of random search \bblcite{Ferreiro2013,Garcia2014,Langdon2011}, to the deterministic algorithms guarantying convergence to the global minimum \bblcite{Evtushenko2009,He2008,Paulavicius2011}.
 
@@
 It  can be achieved by intentional choice of variants in the course of search for the optimal solution cutting  off unpromising search subdomains and investigating only these ones, in which the solution of the  problem can be found.
-The global search algorithm (GSA) is based on this idea \cite{Strongin2000}.
-In  the present work, we tried to combine GSA and a local optimization method (Hooke-Jeeves pattern search method \cite{HookJeeves}) to reduce the number of trials executed.
+The global search algorithm (GSA) is based on this idea \bblcite{Strongin2000}.
+In  the present work, we tried to combine GSA and a local optimization method (Hooke-Jeeves pattern search method \bblcite{HookJeeves}) to reduce the number of trials executed.
 The decision on the run of the local method will be made using a decision tree.
@@
 The local optimization algorithms are intended for determining only one of local extrema within a set  of feasible solutions, in which the objective function takes the extremal (maximum or minimum) value.
-Among the local methods,  the zero-order methods and the gradient methods (of the 1\textsuperscript{st} order and of the  2\textsuperscript{nd} one) are traditionally distinguished \cite{Nocedal}.
-Note that the application of the first-order methods (not to mention the second one) in the problems  with the ``black-box'' type functions is difficult since the problem of numerical estimating the gradient arises here \cite{Kelley}.
+Among the local methods,  the zero-order methods and the gradient methods (of the 1\textsuperscript{st} order and of the  2\textsuperscript{nd} one) are traditionally distinguished \bblcite{Nocedal}.
+Note that the application of the first-order methods (not to mention the second one) in the problems  with the ``black-box'' type functions is difficult since the problem of numerical estimating the gradient arises here \bblcite{Kelley}.
 Therefore, further we will consider only the zero-order methods.
@@
 
-The substantiation of this method of organization of computations see in \cite{Strongin2000,Barkalov2016}.
-The  modifications taking into account the presence of inequality constraints as well as the information on the derivative of the objective function are presented in \cite{Barkalov2002,Gergel1997}.
+The substantiation of this method of organization of computations see in \bblcite{Strongin2000,Barkalov2016}.
+The  modifications taking into account the presence of inequality constraints as well as the information on the derivative of the objective function are presented in \bblcite{Barkalov2002,Gergel1997}.
 
@@
 The Hooke--Jeeves pattern search algorithm belongs to the class of zero-order methods.
-Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{Himmelblau72}.
+Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \bblcite{Himmelblau72}.
 
@@
 	\begin{center}
-		\begin{minipage}[h]{0.8\linewidth}
+		\begin{minipage}[h]{0.7\linewidth}
 			\includegraphics[width=1\linewidth]{figure/fig5.png}
@@
 	\begin{center}
-		\begin{minipage}[h]{0.8\linewidth}
+		\begin{minipage}[h]{0.7\linewidth}
 			\includegraphics[width=1\linewidth]{figure/fig4.png}
@@
 OpenCV is an open source library of the algorithms of computer vision, image processing, and  general-purpose numerical algorithms.
-More details on the decision tree can be found  in \cite{Brahmbhatt2013}.
+More details on the decision tree can be found  in \bblcite{Brahmbhatt2013}.
 
@@
 %Ниже приведены результаты сравнения двух параллельных алгоритмов --- асинхронного GSA и его модификации с использованием decision tree и локального метода Хука-Дживса.
-Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \cite{Sergeyev2006}.
+Численное сравнение проводилось на классах функций Simple и Hard размерности 2, 3, 4 и 5 из \bblcite{Sergeyev2006}.
 Критерием остановки служило попадание точки очередного испытания trial в $\varepsilon$ окрестность истинного глобального минимума.
@@
 Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (AGSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented.
-The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard from \cite{Sergeyev2006}) of dimensionalities 2, 3, 4, and 5.
+The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard from \bblcite{Sergeyev2006}) of dimensionalities 2, 3, 4, and 5.
 These two classes differ in the size of the attraction region of the global minimum point. 
@@
 In many applied optimization problems the dependence of the objective function on some parameters is either linear or unimodal.
-Separating such variables into a special group and solving the problem using a parallel recursive optimization scheme \cite{Barkalov2020_1} can reduce the time for solving the problem by orders of magnitude compared to using a global search for all variables at once.
+Separating such variables into a special group and solving the problem using a parallel recursive optimization scheme \bblcite{Barkalov2020_1} can reduce the time for solving the problem by orders of magnitude compared to using a global search for all variables at once.
 
-\medskip
+\vspace{1em}
+{\it This study was supported by the Russian Science Foundation, project No.\,21-11-00204 .}
 
-\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No.\,21-11-00204 .
-
-%
-% ---- Bibliography ----
-%
-\bibliographystyle{spmpsci}
-\bibliography{bibliography}{}
+\insertbiblioen
 

### 2023-07-27 2023/vestnik.susu/example.tex
minor correction in the abstract

@@
 \begin{abstract}%
-В работе рассматривается решение многомерных задач глобальной оптимизации с использованием деревьев решений для выявления областей притяжения локальных минимумов.
-Целевая функция представлена как «черный ящик», для которой известно, лишь что она удовлетворяет условию Липшица с неизвестной константой.
-Для решения этой задачи применятся алгоритм глобального поиска.
-В работе предлагается метод определения окрестности локальных минимумов целевой функции на основе анализа накопленной поисковой информации.
+В работе рассматривается решение многомерных задач многоэкстремальной оптимизации с использованием деревьев решений для выявления областей притяжения локальных минимумов.
+Целевая функция представлена как «черный ящик»; для нее лишь предполагается, что она удовлетворяет условию Липшица с априори неизвестной константой.
+Для решения задачи многоэкстремальной оптимизации применятся алгоритм глобального поиска.
+Предложена модификация алгоритма, в которой определяются окрестности локальных минимумов целевой функции на основе анализа накопленной поисковой информации.
 Проведение такого анализа с использованием методов машинного обучения позволяет принять решение о запуске локального метода, что может ускорить сходимость алгоритма.

### 2023-07-27 2023/vestnik.susu/example.tex
Поправлена информация об авторах.

@@
 % Сведения об авторах (Ф.И.О.(полностью), степень, звание, кафедра (отдел), вуз, e-mail)
-Первый Александр Васильевич, д.ф.-м.н., профессор, кафедра системного программирования, Южно-Уральский государственный университет (национальный исследовательский университет) (Челябинск, Российская Федерация)
+Баркалов Константин Александрович, д.т.н., доцент, Кафедра математического обеспечения и суперкомпьютерных технологий, Нижегородский государственный университет им.
+Н.И.
+Лобачевского (Н.Новгород, Российская Федерация)
 
-Второй Владимир Григорьевич, к.т.н., доцент, кафедра суперкомпьютеров и квантовой информатики, Московский государственный университет имени М.В.~Ломоносова (Москва, Российская Федерация)
+Лебедев Илья Генадьевич, заведующий лабораторией суперкомпьютерных технологий и высокопроизводительных вычислений,  Нижегородский государственный университет им.
+Н.И.
+Лобачевского (Н.Новгород, Российская Федерация)
+
+Силенко Дмитрий Игоревич, студент, Нижегородский государственный университет им. Н.И. Лобачевского (Н.Новгород, Российская Федерация)
 

### 2023-07-28 2023/ITMO paper/elsarticle-template-num.tex
description of the algorithm

--- 2023-06-14 2023/ITMO paper/elsarticle-template-num.tex
@@
 \label{sec_intro}
-Подбор гиперпараметров - важная и сложная задача.
 
-Много разных фреймворков, у всех есть свои достоинства и недостатки.
+В современных условиях сложные научно и/или промышленные задачи чаще всего решаются численными методами.
+В большинстве таких методов, например, в методах машинного обучения, имеются гиперпараметры, от выбора конкретных значений которых может зависеть эффективность (в заданной метрике) полученного решения при заданных тем или иным способом ограничениях на время расчетов.
+Тот факт, что во многих случаях разница в эффективности получаемого решения при разных значениях гиперпараметров может быть достаточно значительной, привел к возникновению the hyperparameter optimization (HPO) problem, а также исследованиям подходов к ее решению, реализуемых в фреймворках автоматического подбора гиперпараметров.
+С математической точки зрения HPO problem соответствует задаче глобальной оптимизации с фиксированными границах изменения гиперпараметров.
+В некоторых случаях часть гиперпараметров может быть категориальной, то есть принимать значения лишь из некоторого, обычно небольшого, числа вариантов, что приводит к необходимости использовать методы оптимизации, способные решать задачи с дискретными параметрами.
+
+Для конечного пользователя, желающего провести настройку используемого метода, подобрав оптимальное сочетание гиперпараметров, выбор конкретного фреймворка определяется следующими факторами: 1) эффективность реализованных в фреймворке методов, позволяющих решить HPO problem того класса задач, которому принадлежит задача пользователя; 2) простота использования фреймворка с точки зрения описания/формирования задачи подбора оптимального сочетания гиперпараметров, организации процесса поиска оптимума, а также получения и анализа результатов.
+В настоящее время наиболее распространенным способом организации такого интерфейса является использование языка программирования Python, при этом сами фреймворки могут быть написаны, как на этом, так и на других языках.
+
+В данной статье мы представляем фреймворк методов интеллектуальной оптимизации iOpt, ориентированный на решение задач выбора оптимальных значений гиперпараметров различных алгоритмов (включая методы машинного обучения и эвристической оптимизации).
+Статья построена следующим образом.
+В разделе х1 описана алгоритмическая основа методов глобальной оптимизации, реализованных в фреймворке, включая подходы к редукции многомерных задач к одномерным и решению задач с дискретными параметрами.
+В разделе х2 представлено описание фреймворка iOpt.
+В разделе х3 даны результаты численных экспериментов, как демонстрирующих использование фреймворка iOpt, так и сравнивающих его эффективность с другими известными фреймворками на модельных и прикладных задачах.
+В заключении приведены выводы и планы по развитию фреймворка.
 
@@
 
-Простые задачи - метод перебора
+Основная сложность при решении HPO problem состоит в том, что поиск оптимального сочетания гиперпараметров требует для каждого выбранного варианта их значений решения соответствующей исходной задачи, например, задачи машинного обучения, а значит, может быть довольно длительным по времени.
+Таким образом, любой подход к решению HPO problem будет ограничен в числе сочетаний гиперпараметров, которое метод глобальной оптимизации сможет проверить прежде, чем будет исчерпан доступный вычислительный ресурс.
+Конечно, в таких условиях сложно говорить о сходимости метода к глобальному оптимуму или о каких-то гарантиях его отыскания.
+Речь может идти только о возможности получения приемлемого (с точки зрения исследователя) решения.
+Дадим краткую характеристику ряда методов, которые используются для задач настройки гиперпараметров (в зависимости от свойств исходно задачи).
 
-Задачи большой размерности - метаэвристические алгоритмы
+Самым простым способом поиска глобального оптимума является метод полного перебора (поиска по сетке).
+Выбрав число анализируемых вариантов по каждому из гиперпараметров, перебрав все получившиеся сочетания и взяв наилучшее из них в заданной метрике, мы получим оценку оптимума с определенной точностью.
+Метод полного перебора прост в реализации, обладает идеальным параллелизмом, но имеет существенный недостаток – экспоненциальный рост числа узлов сетки с ростом числа гиперпараметров.
+Таким образом, в качестве метода для решения HPO problem полный перебор может использоваться только в «простых» задачах, где малО как число гиперпараметров (один-два), так и время решения исходной задачи при каждом выбранном их сочетании.
 
-Задачи малой размерности - байесовская оптимизация.
+Альтернативой полному перебору при увеличении числа гиперпараметров являются метаэвристические (генетические и аналогичные) алгоритмы.
+Данные алгоритмы широко применяются при отсутствии формульного описания исследуемой модели (модель вида «черный ящик»), что характерно для рассматриваемых HPO problems.
+Метаэвристические алгоритмы слабо зависят от числа гиперпараметров, используют информацию с предыдущих итераций для выполнения текущей, однако в силу заложенной в методы случайности дают гарантию отыскания глобального оптимума только в вероятностном смысле.
 
-Липшицева оптимизация - основные понятия, ссылки на работы. Соотношение с другими подходами. Проблема дискртетных параметров.
+Еще один подход к поиску глобального оптимума -- байесовская оптимизация, также применяемая для решения задач вида «черный ящик».
+Фактически методы байесовской оптимизации строят стохастическую модель оптимизируемой функции.
+Модель итерационно обновляется на основе накапливаемой в процессе поиска оптимума информации, позволяя на каждой очередной итерации оценить наиболее вероятное положение глобального оптимума.
+Методы байесовской оптимизации обладают существенно более высокой эффективностью в сравнении с полным перебором, но, тем не менее, также подвержены влиянию «проклятия размерности», пусть и в меньшей степени.
+
+В известных фреймворках по настройке гиперпараметров широко используется алгоритм Tree-Structured Parzen Estimator (TPE).
+TPE -- это итерационный алгоритм, идейно близкий к методам байесовской оптимизации.
+Перед началом работы алгоритма должна быть задана область поиска в пространстве гиперпараметров и определена оптимизируемая функция.
+В процессе своей работы метод проводит испытания в области поиска, после чего на их основе строит оценки распределения лучших и всех остальных результатов в пространстве гиперпараметров.
+Далее проводится поиск сочетания гиперпараметров, которое дает максимальное ожидаемое улучшение.
+These hyperparameters are then evaluated on the objective function.
+Описанный процесс повторяется до достижения заданного числа итераций или до тех пор, пока не будет достигнуто ограничение по времени.
+
+Хорошие результаты, достигаемые методами байесовской оптимизации и TPE, основаны на априорном предположении, что функция соответствует определенной стохастической модели, например, Gaussian process.
+Однако в глобальной оптимизации существуют и иные предположения о виде функции, которые дают замечательные результаты
+.
+Одним из таких допущений о решаемой задаче является предположение об ограниченности относительных изменений целевой функции.
+В этом случае говорят, что функция удовлетворяет условию Липшица, а решаемая задача называется задачей липшицевой глобальной оптимизации.
+Для класса задач липшицевой глобальной оптимизации разработан целый ряд эффективных алгоритмов [], которые превосходят многие другие методы глобальной оптимизации [] и показали свою применимость ряде приложений [].
+
+Фреймворк iOpt содержит реализацию алгоритмов липшицевой оптимизации, являющихся развитием информационно-статистического алгоритма глобального поиска, детально описанного в []. 
+В частности, методы был адаптирован для решения задач с частично целочисленными параметрами, которые весьма часто встречаются в задачах HPO.
+Результаты экспериментов, приведенные в Section XX, показывают, что в задачах HPO фреймворк iOpt не уступает известным фреймворкам аналогичного назначения (таким как Optuna и HyperOpt).
 
@@
 \subsection{Проблема дискретных параметров}
+\label{sec_discr}
 
@@
 имеющую, вообще говоря, разрывы в целочисленных точках $x_i = i, 1\leq i \leq S-1$.
-Поэтому значения  $z_i = f(x_i)$ в этих точках будем считать неопределенным, а значения индекса -- равным 0, т.е. $\nu(x_i) = 0$.
+Поэтому значения  $z_i = f(x_i)$ в этих точках будут считаться неопределенными и не будут использоваться в алгоритме.
+%, а значения индекса -- равным 0, т.е. $\nu(x_i) = 0$.
 
@@
 
-Test reference \cite{Strongin2000,Barkalov2021}.
-
 \section{iOpt framework} 
 \label{sec_iOpt}
+
+Дать характеристику фреймворка - язык разработки, используемые технологии, ориентация на какие задачи, и т.п.
 
@@
 
+Алгоримтическим ядром фреймворка iOpt является информационно-статистический алгоритм глобального поиска для решения задач вида (\ref{f_func}), (\ref{f_D}). 
+Данный алгоритм предполагает построение последовательности точек $y^i$, в которых проводятся поисковые испытания, т.е.
+вычисляются значения objective function $z^i = \varphi(y^i)$.
+В соответствии с используемой схемой редукции размерности проведение испытания подразумевает вычисление $y^i=y(x^i))$, поэтому результатом испытания будет являться набор значений $(x^i, y^i=y(x^i), z^i = \varphi(y^i))$.
+Поисковую информацию, накопленную после проведения $k$ испытаний, удобно представлять в виде множества, в котором соответствующие наборы перенумерованы (нижним индексом) в порядке возрастания $x$, т.е.
+в виде множества
+\begin{equation}\label{omega}
+\Omega_k = \left\{  (x_i, y_i=y(x_i), z_i = \varphi(y_i)): 1 \leq i \leq k  \right\},	
+\end{equation}
+где $x_1 < x_2 < ... < x_k$.
+
+Основная идея алгоритма глобального поиска состоит в следующем.
+Пусть в процессе поиска проведено $k \geq 2$ испытаний and the search information $\Omega_k$ from (\ref{omega}) is obtained.
+Для выбора точки очередного испытания для каждого поискового интервала $(x_{i-1},x_i), 1<i\leq k$, вычисляется его характеристика в соответствии с формулой
+\begin{equation}\label{R}
+R(i) = \Delta_i + \frac{(z_i-z_{i-1})^2}{(r\mu)^2\Delta_i}-2\frac{z_{i-1}+z_i}{r\mu}.	
+\end{equation}
+Здесь $r>1$ -- параметр алгоритма, $\Delta_i=(x_i-x_{i-1})^{1/N}$.
+Значение $\mu$ является нижней оценкой константы Гельдера целевой функции, вычисленной на основе поисковой информации из (\ref{omega}) по формуле
+\begin{equation}\label{mu}
+\mu = \max_{1<i\leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
+\end{equation}
+причем в случае, если формула (\ref{mu}) выдает значение $\mu=0$, то используется значение $\mu=1$.
+
+Новое испытание проводится в точке $x^{k+1}$, принадлежащей интервалу с наибольшей характеристикой.
+Если правая точка этого интервала имеет номер $t$ в множестве $\Omega_k$, то формула для вычисления $x^{k+1}$ будет выглядеть следующим образом:
+\begin{equation}\label{xk1}
+x^{k+1} = \frac{x_t+x_{t-1}}{2}- \frac{\mathrm{sign}(z_t-z_{t-1})}{2r} \left(\frac{\left|z_t-z_{t-1}\right|}{\mu}\right)^N.   
+\end{equation}
+
+Algorithm \ref{alg_GSA} shows the pseudo code of the Global Search Algorithm.
+Входными данными алгоритма являются objective function $\varphi(y)$  и область поиска $D$ из (\ref{f_func}) and (\ref{f_D}), параметр алгоритма $r>1$, точность поиска минимума $\epsilon > 0$, максимально допустимое число испытаний $K_{max}$.
+Выходным параметром алгоритма является оценка точки глобального минимума $y_{min}$.
+Дополнительными выходными параметрами могут быть значение функции в точке минимума $\varphi(y_{min})$ и число проведенных испытаний $k$.
+
+\begin{algorithm}
+\LinesNumbered
+ \KwIn{$\varphi(y), D, r, \epsilon, K_{max}$}
+ \KwOut{$y_{min}$}
+ Initialize $k=2$, $\Omega_k= \left\{ (x_1=0, y_1=y(x_1), z_1=\varphi(y_1)), (x_2=1, y_2=y(x_2), z_2=\varphi(y_2)) \right\}$\\
+ \While{ $\Delta_t \geq \epsilon$ and  $k \leq K_{max}$}{
+  Вычислить оценку H\"older constant $\mu$ в соответствии с (\ref{mu})\\
+	\For{$i=1$ \KwTo $k$}{
+	Вычислить характеристику $R(i)$ в соответствии с (\ref{R})\\
+	}
+	Определить номер $t$, для которого $R(t) = \max \left\{ R(i), 1 \leq i \leq k \right\}$
+
+	Вычислить точку нового испытания $x^{k+1}$ в соответствии с (\ref{xk1})
+
+	Провести новое испытание в точке $x^{k+1}$
+
+	Добавить $(x^{k+1}, y^{k+1}, z^{k+1})$ в множество $\Omega_k$
+
+  $k = k + 1$\\
+ }
+ $l = \arg \min \left\{ z_i, 1 < i \leq k \right\}$\\
+ $y_{min} = y_l$, \\ 
+ \KwRet{$y_{min}$}
+ \caption{Global search}\label{alg_GSA}
+\end{algorithm}
+
+Теория сходимости алгоритма, а также некоторые его модификации представлены в \cite{Strongin2000}.
+
 \subsection{GSA for the problems with categorical parameters}
@@
 
-А нужно ли тут отдельное описание?
+Рассмотрим случай, когда в настраиваемом алгоритме имеются как непрерывные, так и категориальные параметрами. Характерными примером является support vector machine.
+
+Дать описание параметров svm - непрерывные и категориальные параметры.
+
+В этом случае категориальным параметрам можно поставить в соответствие целочисленные значения , и свести сзадачу к постановке (\ref{problem_is1}).
+
+Так как решение задач с частично целочисленными параметрами в соответсвии со схемой, изложенной в Section \ref{sec_discr}, предполагает решение задачи (\ref{problem_is1}) с разрывной целевой функцией, требуется модификация алгоритма, учитывающая эту особенность.
+
+С этой целью введем классификацию точек из области поиска с помощью специального индекса $\nu_i=\nu(x_i)$.
+Индексы целочисленных точек $x_i = i, 0\leq i \leq S$, будем полагать равным нулю, т.е.
+$\nu(x_i) = 0$.
+Для всех остальных значений $x\in(i,i+1),  0 \leq i < S$, значение индекса будем считать равным единице, т.е.
+$\nu(x) = 1$.
+Тогда результатом испытания в некоторой точке $x^j\in[0,S]$ будет являться набор значений
+$(x^j, \nu^j=\nu(x^j), y^j=y(x^j), z^j = \varphi(y^j))$. Причем в случае $\nu^j=0$ значение $z^j$ будет считаться неопределенным.
+
+Общая вычислительная схема, изложенная в виде Алгоритма \ref{alg_GSA}, останется прежней. Модификации будут состоять в следующем.
+Модификации будут состоять в следующем.
+
+1.
+На этапе инициализации в множество $\Omega_k$, содержащее поисковую информацию, заносятся значения $(x^i = i, \nu^i=0, y^i=y(x^i), z^i)$, $ i, 0\leq i \leq S$, т.е.
+заносятся граничные точки интервалов $(x_{i-1},x_i), 1<i\leq S$.
+После этого проводятся испытания в произвольных внутренних точках (например, в серединах) интервалов $(x_{i-1},x_i), 1<i\leq S$; результаты испытаний также добавляются в поисковую информацию. 
+Таким образом, после инициализации множество $\Omega_k$ будет содержать $k=2S+1$ элемент.
+
+2.
+Вычисление оценки константы Гельдера $\mu$ в соответствии с формулой (\ref{mu}) должно выполняться только для  интервалов, для которых $\nu(x_i) = \nu(x_{i-1}) = 1$, т.е.
+когда значения $z_i$ and $z_{i-1}$ определены.
+
+3. Вычисление характеристики поискового интервала $(x_{i-1},x_i), 1<i\leq k$, проводится по следующим правилам:
+\begin{gather}\label{R_int}
+R(i) = \Delta_i + \frac{(z_i-z_{i-1})^2}{(r\mu)^2\Delta_i}-2\frac{z_{i-1}+z_i}{r\mu}, \; \mathrm{if} \; \nu(x_i) = \nu(x_{i-1}) = 1, \nonumber \\ 
+R(i) = 2\Delta_i-4\frac{z_i}{r\mu}, \; \mathrm{if} \;  \nu(x_i) = 1, \nu(x_{i-1}) = 0, \nonumber \\ 
+R(i) = 2\Delta_i-4\frac{z_{i-1}}{r\mu}, \; \mathrm{if} \;  \nu(x_{i-1}) = 1, \nu(x_{i}) = 0. \nonumber
+\end{gather}
+Отметим, что в процессе работы алгоритма не будет возникать интервалов, у которых обе граничные точки имеют неопределенные значения функции, т.е. $\nu(x_i) = \nu(x_{i-1}) = 0$.
+
+4. Точка нового испытания $x^{k+1}$ будет вычисляться следующим образом:
+\begin{gather}\label{xk1_int}
+x^{k+1} = \frac{x_t+x_{t-1}}{2}- \frac{\mathrm{sign}(z_t-z_{t-1})}{2r} \left(\frac{\left|z_t-z_{t-1}\right|}{\mu}\right)^N, \; \mathrm{if} \; \nu(x_i) = \nu(x_{i-1}) = 1, \nonumber \\    
+x^{k+1} = \frac{x_t+x_{t-1}}{2} , \; \mathrm{if} \; \nu(x_i) \neq \nu(x_{i-1}). \nonumber
+\end{gather}
+Здесь $t$ есть номер поискового интервала, имеющего максимальную характеристику.
 

### 2023-07-31 2023/ITMO paper/elsarticle-template-num.tex
add some references

@@
 В этом случае говорят, что функция удовлетворяет условию Липшица, а решаемая задача называется задачей липшицевой глобальной оптимизации.
-Для класса задач липшицевой глобальной оптимизации разработан целый ряд эффективных алгоритмов [], которые превосходят многие другие методы глобальной оптимизации [] и показали свою применимость ряде приложений [].
+Для класса задач липшицевой глобальной оптимизации разработан целый ряд эффективных алгоритмов \cite{Jones2021,Paulavicius2020,Strongin2020,Sergeyev2017,PaulaviciusZilinskas2014}, которые превосходят многие другие методы глобальной оптимизации \cite{Sergeyev2018} и показали свою применимость ряде приложений \cite{Gubaydullin2021}.
 
-Фреймворк iOpt содержит реализацию алгоритмов липшицевой оптимизации, являющихся развитием информационно-статистического алгоритма глобального поиска, детально описанного в []. 
-В частности, методы был адаптирован для решения задач с частично целочисленными параметрами, которые весьма часто встречаются в задачах HPO.
-Результаты экспериментов, приведенные в Section XX, показывают, что в задачах HPO фреймворк iOpt не уступает известным фреймворкам аналогичного назначения (таким как Optuna и HyperOpt).
+Фреймворк iOpt содержит реализацию алгоритмов липшицевой оптимизации, являющихся развитием информационно-статистического алгоритма глобального поиска, детально описанного в \cite{Strongin2000,Sergeyev2013}.
+В частности, методы были адаптированы для решения задач с частично целочисленными параметрами, которые весьма часто встречаются в задачах HPO.
+Результаты экспериментов, приведенные в Section XX, показывают, что в задачах HPO фреймворк iOpt не уступает известным фреймворкам аналогичного назначения (таким как Optuna \cite{optuna} and HyperOpt \cite{hyperopt}).
 
@@
 
+Еще раз отметим, что характеристика $R(i)$ количественно оценивает возможность нахождения точки $x^*$, являющейся inverse image of the global optimizer $y* = y(x^*)$, в пределах рассматриваемого интервала $(x_{i-1},x_i)$.
+
 \section{Numerical experiments}

### 2023-08-01 2023/ITMO paper/elsarticle-template-num.tex
add two pictures

@@
 В этом случае говорят, что функция удовлетворяет условию Липшица, а решаемая задача называется задачей липшицевой глобальной оптимизации.
-Для класса задач липшицевой глобальной оптимизации разработан целый ряд эффективных алгоритмов \cite{Jones2021,Paulavicius2020,Strongin2020,Sergeyev2017,PaulaviciusZilinskas2014}, которые превосходят многие другие методы глобальной оптимизации \cite{Sergeyev2018} и показали свою применимость ряде приложений \cite{Gubaydullin2021}.
+Для класса задач липшицевой глобальной оптимизации разработан целый ряд эффективных алгоритмов \cite{Jones2021,Paulavicius2020,Strongin2020,Sergeyev2017,PaulaviciusZilinskas2014}, которые превосходят многие другие методы глобальной оптимизации \cite{Sergeyev2018} и показали свою применимость ряде приложений \cite{Kvasov2008,CANDELIERI2019,Gubaydullin2021}.
 
@@
 В частности, методы были адаптированы для решения задач с частично целочисленными параметрами, которые весьма часто встречаются в задачах HPO.
-Результаты экспериментов, приведенные в Section XX, показывают, что в задачах HPO фреймворк iOpt не уступает известным фреймворкам аналогичного назначения (таким как Optuna \cite{optuna} and HyperOpt \cite{hyperopt}).
+Результаты экспериментов, приведенные в Section \ref{sec_exp}, показывают, что в задачах HPO фреймворк iOpt не уступает известным фреймворкам аналогичного назначения (таким как Optuna \cite{optuna} and HyperOpt \cite{hyperopt}).
 
@@
 \end{equation}
-Здесь $ \left\| \cdot \right\|$ обозначает евклидову норму, а константа $L$ является априори не изветной и подлежит оценке в процессе решения. 
+Здесь $ \left\| \cdot \right\|$ обозначает евклидову норму, а константа $L$ является априори не известной и подлежит оценке в процессе решения. 
 %Такие задачи называются задачами липшицевой глобальной оптимизации with box-constraints.
@@
 
-В силу данных неравенств значения функции в точках интервала $[x', x'']$ должны располагаться внутри области, ограниченной прямыми, проходящими через точки $(x', f(x'))$ и $(x'', f(x''))$ с наклоном $L$ и $-L$ (см.
-рис.
+В силу данных неравенств значения функции в точках интервала $[x', x'']$ должны располагаться внутри области, ограниченной прямыми, проходящими через точки $(x', f(x'))$ и $(x'', f(x''))$ с наклоном $L$ и $-L$ (see Fig.
 ).
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{Fig1.pdf}
+\caption{Значения функции $f(x)$ должны находиться внутри области, выделенной цветом} \label{fig1}
+\end{figure}
 
@@
 
-Методы липшицевой глобальной оптимизации используют эту идею в своих вычислительных правилах; на ней также основано и доказательство их сходимости (see, e.g., ).
+Методы липшицевой глобальной оптимизации используют эту идею в своих вычислительных правилах; на ней также основано и доказательство их сходимости (see, e.g.,
+\cite{Jones2021,PaulaviciusZilinskas2014,Sergeyev2013,Evtushenko2013}).
 
@@
 
-Известным подходом к решению многомерной задачи (\ref{f_func}) является ее редукция к одномерной задаче с последующим применением эффективных одномерных алгоритмов глобальной оптимизации (see []).
+Известным подходом к решению многомерной задачи (\ref{f_func}) является ее редукция к одномерной задаче с последующим применением эффективных одномерных алгоритмов глобальной оптимизации (see \cite{Strongin2000,Sergeyev2013}).
 Здесь хорошо зарекомендовал себя подход, в котором размерность задачи редуцируется при помощи кривой Пеано–Гильберта $y(x), x \in [0, 1]$.
@@
 
-При этом известно (see []), что если многомерная функция $\varphi(y), \; y \in D,$  удовлетворяет условию Липшица с константой $L$, то редуцированная одномерная функция $\varphi(y(x)), \; \in [0,1]$ удовлетворяет the H{\"o}lder condition
+При этом известно (see \cite{Strongin2000}), что если многомерная функция $\varphi(y), \; y \in D,$  удовлетворяет условию Липшица с константой $L$, то редуцированная одномерная функция $\varphi(y(x)), \; \in [0,1]$ удовлетворяет the H{\"o}lder condition
 \[
@@
 Замечание 1. Для редуцированной одномерной функции $\varphi(y(x))$  не будет выполняться условие Липшица. Однако многие одномерные алгоритмы липшицевой оптимизации
-могут быть обобщены на случай минимизации гeльдеровых функций (see []). Конкретный пример такого обобщения приведен в subsection \ref{sec_GSA}.
+могут быть обобщены на случай минимизации гeльдеровых функций. Конкретный пример такого обобщения приведен в subsection \ref{sec_GSA}.
+
+\begin{figure}
+\center
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=0.9\linewidth]{fig2a.jpg} }
+\end{minipage}
+\begin{minipage}{0.45\linewidth}
+\center{\includegraphics[width=0.9\linewidth]{fig2b.jpg} }
+\end{minipage}
+\caption{Evolvents $y_m(x)$ with $m=4$.}\label{fig:Peano}
+\end{figure}
 
@@
 Теоретическая кривая Пеано-Гильберта $y(x)$ определяется как предельный объект.
-В численных алгоритмах применяются развертки $y_m(x)$, аппроксимирующие истинную кривую Пеано–Гильберта c заданным уровнем точности $2^{-m}$, зависящим от требуемой точности поиска.
-Эффективные схемы вычисления нескольких видов разверток подробно описаны в \cite{Sergeyev2013}.
-%пример развертки - рисунок с функцией
+В численных алгоритмах применяются evolvents $y_m(x)$, аппроксимирующие истинную curve c заданным уровнем плотности $2^{-m}$, зависящим от требуемой точности поиска.
+Эффективные схемы вычисления нескольких видов evolvents подробно описаны в \cite{Sergeyev2013}.
+Для иллюстрации на Fig.~\ref{fig:Peano} приведены графики, соответствующие разверткам $y_m(x)$ с плотностью $m=4$ для размерностей $N=2$ и $N=3$.
 
@@
 для них сложнее построить оценки оптимума по сравнению с задачами с непрерывными параметрами.
-
 Методам решения задач со смешанными параметрами посвящено много публикаций (см., например, обзоры \cite{Burer2012,Boukouvala2016}). 
-Известные детерминированные методы решения задач данного класса ориентированы на решение линейных или выпуклых задач.
-Многоэкстремальные задачи предлагается решать с помощью метаэвристических алгоритмов.
+Известные детерминированные методы решения задач данного класса ориентированы на решение линейных или выпуклых задач \cite{Lee2012}.
+Сложные многоэкстремальные задачи предлагается решать с помощью эвристических подходов \cite{Belotti2013}.
 
 Нами был предложен и реализован новый детерминированный подход к решению задач со смешанными параметрами; приведем здесь его краткое описание. 
-
 Пусть  целевая функция задачи зависит от двух векторов параметров: вектора $y$, принадлежащего гиперинтервалу $D$, и вектора $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е.
@@
 Такие конечные наборы значений могут характеризовать множество возможных сочетаний категориальных гиперпараметров исследуемого алгоритма машинного обучения. 
-%Целочисленные параметры = непрерывные параметры с округлением, добавить комментарий об этом.
+Характерными примером является support vector machine, влияние на качество работы которого оказывают непрерывные параметр регуляризации $C$ и коэффициент ядра $\gamma$.
+При этом тип функции ядря выбирается из конечного множества вариантов (например, полиномиальная функция, радиальная базисная функция, сигмоид и т.д.).
+%Целочисленные параметры = непрерывные параметры с округлением, добавить комментарий об этом?
 
@@
 
-Рассмотрим случай, когда в настраиваемом алгоритме имеются как непрерывные, так и категориальные параметрами. Характерными примером является support vector machine.
-
-Дать описание параметров svm - непрерывные и категориальные параметры.
-
-В этом случае категориальным параметрам можно поставить в соответствие целочисленные значения , и свести сзадачу к постановке (\ref{problem_is1}).
-
+Рассмотрим случай, когда в настраиваемом алгоритме имеются как непрерывные, так и категориальные параметрами. 
+В этом случае категориальным параметрам можно поставить в соответствие целочисленные значения, и свести сзадачу к постановке (\ref{problem_is1}).
 Так как решение задач с частично целочисленными параметрами в соответсвии со схемой, изложенной в Section \ref{sec_discr}, предполагает решение задачи (\ref{problem_is1}) с разрывной целевой функцией, требуется модификация алгоритма, учитывающая эту особенность.

### 2023-08-02 2023/ITMO paper/elsarticle-template-num.tex
title, abstract and highlights

@@
 
-\title{Title of the paper}
+\title{On hyperparameter tuning through Lipschitz global optimization} % and dimensionality reduction}?
+%% О настройке гиперпараметров с помощью липшицевой глобальной оптимизации
 
@@
 
+На качество работы методов машинного обучения существенное влияние оказывают их гиперпараметры, при этом оценка критерия качества является трудоемкой операцией.
+Поэтому актуальной является задача разработки ителлектуальных методов для выбора оптимальных значений гиперпараметров, которые требуют малого числа поисковых испытаний.
+В этой статье мы предлагаем новый подход к настройке гиперпараметров, основанный на идеях Lipschitz global optimization. 
+В рамках данного подхода решение задач с несколькими параметрами сводится к решению эквивалентных им одномерных задач.
+Соответствующая редукция основана на использовании space-filling curves (Peano curves).
+Указанные подходы реализованы в open-source фреймворке методов интеллектуальной оптимизации iOpt.
+Для демонстрации преимуществ iOpt проведено его сравнение с известными фреймворками Optuna и HyperOpt при настройке гиперпараметров различных методов машинного обучения на представительном наборе датасетов.
+Результаты показывают, что методы Lipschitz global optimization обеспечивают сравнимые по качеству результаты за существенно меньшее время по сравнению с известными алгоритмами настройки гиперпараметров.
+
 \end{abstract}
@@
 \begin{highlights}
-\item Research highlight 1
-\item Research highlight 2
+\item Для настройки гиперпараметров разработан фреймворк методов липшицевой оптимизации
+\item Для решения многомерных задач используется редукция размерности на оснвое кривых Пеано
+\item Предложена схема сведения задач, часть параметров которых является категориальными, к одной задаче с непрерывными параметрами 
+\item Разработанные методы (в отличие от известных методов настройки гиперпараметров) являются детерминированными, что обеспечивает стабильный результат при разных запусках алгоритмов
 \end{highlights}
@@
 %% keywords here, in the form: keyword \sep keyword
+Lipschitz optimization \sep Multiextremal function \sep Hyperparameter tuning
 

### 2023-08-15 2023/ITMO paper/elsarticle-template-num.tex
minor corrections before translation of the first part

@@
 
-В современных условиях сложные научно и/или промышленные задачи чаще всего решаются численными методами.
+В настоящее время сложные научно и/или промышленные задачи чаще всего решаются численными методами.
 В большинстве таких методов, например, в методах машинного обучения, имеются гиперпараметры, от выбора конкретных значений которых может зависеть эффективность (в заданной метрике) полученного решения при заданных тем или иным способом ограничениях на время расчетов.
@@
 Статья построена следующим образом.
-В разделе х1 описана алгоритмическая основа методов глобальной оптимизации, реализованных в фреймворке, включая подходы к редукции многомерных задач к одномерным и решению задач с дискретными параметрами.
-В разделе х2 представлено описание фреймворка iOpt.
-В разделе х3 даны результаты численных экспериментов, как демонстрирующих использование фреймворка iOpt, так и сравнивающих его эффективность с другими известными фреймворками на модельных и прикладных задачах.
+В начале приведен обзор известных методов, используемых для настройки гиперпараметров.
+В section \ref{sec_lip} приведена математическая постановка задачи, описаны одходы к редукции многомерных задач к одномерным и решению задач с дискретными параметрами.
+В section \ref{sec_iOpt} описана алгоритмическая основа методов глобальной оптимизации, реализованных во фреймворке iOpt.
+В section \ref{sec_exp} даны результаты численных экспериментов, проведенных для сравнения эффективности фреймворка iOpt с другими известными фреймворками на модельных и прикладных задачах.
 В заключении приведены выводы и планы по развитию фреймворка.
@@
 Хорошие результаты, достигаемые методами байесовской оптимизации и TPE, основаны на априорном предположении, что функция соответствует определенной стохастической модели, например, Gaussian process.
-Однако в глобальной оптимизации существуют и иные предположения о виде функции, которые дают замечательные результаты
-.
+Однако в глобальной оптимизации существуют и иные предположения о виде функции, которые дают замечательные результаты.
+
 Одним из таких допущений о решаемой задаче является предположение об ограниченности относительных изменений целевой функции.
@@
 
-Дать характеристику фреймворка - язык разработки, используемые технологии, ориентация на какие задачи, и т.п.
+Фреймворк методов интеллектуальной оптимизации iOpt разработан в рамках выполенния проекта ``Strong Artificial Intelligence in Industry''.
+Он предназначен для выбора оптимальных (в заданной метрике) значений параметров сложных объектов и процессов, например, методов искусственного интеллекта и машинного обучения.
+Фреймворк позволяет проводить точную настройку параметров моделей и методов, используемых в прикладных исследованиях в различных научных областях.
+Фреймворк iOpt разработан на языке программирования Python с использованием его стандартной библиотеки.
+Для его использования необходим установленный интерпретатор языка Pyhon версии не ниже 3.8.
+Фреймворк является open source и доступен по адресу https://github.com/aimclub/iOpt.
+
+Фреймворк iOpt может использоваться при создании специализированных систем поддержки принятия решений для решения задач выбора оптимальных параметров сложных моделей или методов.
+Область применения фреймворка фокусируется на задачах настройки гиперпараметров, поскольку при решении задач этого класса возникают проблемы многоэкстремальности критерия оптимизации и отсутствия формульного описания исследуемой модели (модель вида «черный ящик»).
 

### 2023-08-21 2023/ITMO paper/elsarticle-template-num.tex
add some references

@@
 В настоящее время сложные научно и/или промышленные задачи чаще всего решаются численными методами.
-В большинстве таких методов, например, в методах машинного обучения, имеются гиперпараметры, от выбора конкретных значений которых может зависеть эффективность (в заданной метрике) полученного решения при заданных тем или иным способом ограничениях на время расчетов.
-Тот факт, что во многих случаях разница в эффективности получаемого решения при разных значениях гиперпараметров может быть достаточно значительной, привел к возникновению the hyperparameter optimization (HPO) problem, а также исследованиям подходов к ее решению, реализуемых в фреймворках автоматического подбора гиперпараметров.
+В большинстве таких методов, например, в методах машинного обучения, имеются гиперпараметры, от выбора конкретных значений которых может зависеть эффективность (в заданной метрике) полученного решения при заданных тем или иным способом ограничениях на время расчетов \cite{Hutter2019}.
+Тот факт, что во многих случаях разница в эффективности получаемого решения при разных значениях гиперпараметров может быть достаточно значительной, привел к возникновению the hyperparameter optimization (HPO) problem, а также исследованиям подходов к ее решению, реализуемых в фреймворках автоматического подбора гиперпараметров \cite{Tune, optuna, hyperopt,Sherpa}.
 С математической точки зрения HPO problem соответствует задаче глобальной оптимизации с фиксированными границах изменения гиперпараметров.
@@
 
-Самым простым способом поиска глобального оптимума является метод полного перебора (поиска по сетке).
-Выбрав число анализируемых вариантов по каждому из гиперпараметров, перебрав все получившиеся сочетания и взяв наилучшее из них в заданной метрике, мы получим оценку оптимума с определенной точностью.
+Самым простым способом поиска глобального оптимума является метод полного перебора (поиска по равномерной \cite{Bao2006} или случайной \cite{Bergstra2012} сетке).
+Выбрав число анализируемых вариантов по каждому из гиперпараметров, перебрав все получившиеся сочетания и взяв наилучшее из них в заданной метрике, мы получим оценку оптимума с определенной точностью \cite{Nevendra2022}.
 Метод полного перебора прост в реализации, обладает идеальным параллелизмом, но имеет существенный недостаток – экспоненциальный рост числа узлов сетки с ростом числа гиперпараметров.
@@
 
-Альтернативой полному перебору при увеличении числа гиперпараметров являются метаэвристические (генетические и аналогичные) алгоритмы.
-Данные алгоритмы широко применяются при отсутствии формульного описания исследуемой модели (модель вида «черный ящик»), что характерно для рассматриваемых HPO problems.
+Альтернативой полному перебору при увеличении числа гиперпараметров являются метаэвристические (генетические и аналогичные) алгоритмы \cite{Opara2019}.
+Данные алгоритмы широко применяются при отсутствии формульного описания исследуемой модели (модель вида «черный ящик»), что характерно для рассматриваемых HPO problems \cite{Zhou2021,Yang2022}.
 Метаэвристические алгоритмы слабо зависят от числа гиперпараметров, используют информацию с предыдущих итераций для выполнения текущей, однако в силу заложенной в методы случайности дают гарантию отыскания глобального оптимума только в вероятностном смысле.
 
-Еще один подход к поиску глобального оптимума -- байесовская оптимизация, также применяемая для решения задач вида «черный ящик».
+Еще один подход к поиску глобального оптимума -- байесовская оптимизация, также применяемая для решения задач вида «черный ящик» \cite{Frazier2018,Archetti2019}.
 Фактически методы байесовской оптимизации строят стохастическую модель оптимизируемой функции.
@@
 
-В известных фреймворках по настройке гиперпараметров широко используется алгоритм Tree-Structured Parzen Estimator (TPE).
+В известных фреймворках по настройке гиперпараметров широко используется алгоритм Tree-Structured Parzen Estimator (TPE) \cite{hyperopt,NIPS2011}.
 TPE -- это итерационный алгоритм, идейно близкий к методам байесовской оптимизации.
@@
 
-Хорошие результаты, достигаемые методами байесовской оптимизации и TPE, основаны на априорном предположении, что функция соответствует определенной стохастической модели, например, Gaussian process.
+Хорошие результаты, достигаемые методами байесовской оптимизации  \cite{Joy2020} и TPE \cite{Watanabe2022a,Watanabe2022b}, основаны на априорном предположении, что функция соответствует определенной стохастической модели, например, Gaussian process \cite{Rasmussen2005}.
 Однако в глобальной оптимизации существуют и иные предположения о виде функции, которые дают замечательные результаты.
@@
 вычисление значения функции в точке допустимой области) требует значительных вычислительных ресурсов.
-Такая постановка проблемы полностью соответствует задаче настройки гиперпараметров методов машинного обучения (see ).
+Такая постановка проблемы полностью соответствует задаче настройки гиперпараметров методов машинного обучения \cite{Joy2020,Wang2021}.
 
@@
 В силу данных неравенств значения функции в точках интервала $[x', x'']$ должны располагаться внутри области, ограниченной прямыми, проходящими через точки $(x', f(x'))$ и $(x'', f(x''))$ с наклоном $L$ и $-L$ (see Fig.
-).
+\ref{fig1}).
 

### 2023-08-22 2023/ITMO paper/elsarticle-template-num.tex
add table

@@
 \usepackage[ruled]{algorithm2e}
+\usepackage{multirow}
 
@@
 
+\begin{table}[]
+\caption{\label{tab:1}14 datasets used in the experiments.}
+
+\begin{tabular}{lllllll}
+\hline
+Name                                                                               & Num   & Attributes & Int & Float & Categorical & Classes \\ \hline
+Bank Marketing{[}5{]}                                                              & 45211 & 20         & 6   & 4     & 10          & 2       \\
+Wilt{[}15{]}                                                                       & 4889  & 5          & 0   & 5     & 0           & 2       \\
+\begin{tabular}[c]{@{}l@{}}Banknote \\ authentication{[}9{]}\end{tabular}          & 1372  & 4          & 0   & 4     & 0           & 2       \\
+\begin{tabular}[c]{@{}l@{}}Credit \\ Approval{[}11{]}\end{tabular}                 & 690   & 15         & 0   & 6     & 9           & 2       \\
+\begin{tabular}[c]{@{}l@{}}Breast Cancer \\ Wisconsin{[}2{]}\end{tabular}          & 569   & 9          & 9   & 0     & 0           & 2       \\
+Parkinsons{[}13{]}                                                                 & 197   & 22         & 0   & 22    & 0           & 2       \\
+\begin{tabular}[c]{@{}l@{}}Balance Scale \\ Weight \& Distance{[}8{]}\end{tabular} & 625   & 4          & 4   & 0     & 0           & 3       \\
+Car Evaluation {[}10{]}                                                            & 1728  & 6          & 0   & 0     & 6           & 4       \\
+\begin{tabular}[c]{@{}l@{}}Statlog \\ (Image Segmentation){[}7{]}\end{tabular}     & 2310  & 19         & 1   & 18    & 0           & 7       \\
+Zoo{[}19{]}                                                                        & 101   & 16         & 16  & 0     & 0           & 7       \\
+\begin{tabular}[c]{@{}l@{}}Protein \\ Localization \\ Sites (Ecoli){[}12{]}\end{tabular}     & 336  & 7  & 0  & 7 & 0 & 8  \\
+CNAE-9{[}3{]}                                                                      & 1080  & 856        & 856 & 0     & 0           & 9       \\
+\begin{tabular}[c]{@{}l@{}}Optical Recognition\\  of Handwritten Digits {[}4{]}\end{tabular} & 5620 & 64 & 64 & 0 & 0 & 10 \\
+\begin{tabular}[c]{@{}l@{}}Semeion Handwritten \\ Digit{[}6{]}\end{tabular}        & 1593  & 256        & 256 & 0     & 0           & 10      \\
+\begin{tabular}[c]{@{}l@{}}Turkiye Student \\ Evaluation{[}14{]}\end{tabular}      & 5820  & 32         & 32  & 0     & 0           & 13      \\ \hline
+\end{tabular}
+\end{table}
+
 Metrics
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+\begin{table}[]
+\caption{Настраиваемые гиперпараметры для алгоритмов машинного обучения}
+\label{tab:2}
+\begin{tabular}{llll}
+\hline
+Метод                    & Параметры            & Тип параметра & Диапазон параметра                                                     \\ \hline
+\multirow{7}{*}{XGBoost} & n\_estimators        & Int           & {[}10, 200{]}                                                          \\
+                         & max\_depth           & Int           & {[}5, 20{]}                                                            \\
+                         & min\_child\_weight   & Int           & {[}1, 10{]}                                                            \\
+                         & Gamma                & Float         & {[}0.01, 0.6{]}                                                        \\
+                         & Subsample            & Float         & {[}0.05, 0.95{]}                                                       \\
+                         & colsample\_bytree    & Float         & {[}0.05, 0.95{]}                                                       \\
+                         & learning\_rate       & Float         & {[}0.001, 0.1{]}                                                       \\ \hline
+\multirow{3}{*}{\begin{tabular}[c]{@{}l@{}}Support \\ Vector   \\ Classification \\ (SVC)\end{tabular}} &
+  Gamma &
+  Float &
+  {[}$10^{-9}$, $10^{-6}${]} \\
+                         & C                    & Int           & {[}$10^1$,$10^{10}${]}                     \\
+                         & Kernel               & Categorial    & \begin{tabular}[c]{@{}l@{}}\{‘poly’, ‘rbf’,\\ ‘sigmoid’\}\end{tabular} \\ \hline
+\multirow{4}{*}{\begin{tabular}[c]{@{}l@{}}Multi-layer   \\ Perceptron \\ (MLP)\end{tabular}} &
+  Activation &
+  Categorial &
+  \begin{tabular}[c]{@{}l@{}}\{'identity', \\ 'logistic', \\ 'tanh', 'relu'\}\end{tabular} \\
+                         & hidden\_layer\_sizes & Int           & {[}2, 30{]}                                                            \\
+                         & max\_iter            & Int           & {[}100, 700{]}                                                         \\
+                         & Solver               & Categorial    & \begin{tabular}[c]{@{}l@{}}\{'lbfgs', 'sgd', \\ 'adam'\}\end{tabular}  \\ \hline
+\end{tabular}
+\end{table}
 
@@
 Third result
+
+\begin{table}[]
+\caption{Время поиска значений гиперпараметров при однократном запуске алгоритма оптимизации.}
+\label{tab:3}
+\begin{tabular}{lllllll}
+\hline
+Dataset &
+  \begin{tabular}[c]{@{}l@{}}Optuna \\ random\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Optuna\\ tpe\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Optuna \\ cmaes\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Optuna \\ nsgaii\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Hyperopt\\ tpe\end{tabular} &
+  iOpt \\ \hline
+Balance                                                        & 10,77   & 9,15    & 9,02    & 10,60   & 10,29   & 7,67    \\
+\begin{tabular}[c]{@{}l@{}}Bank\\ Marketing\end{tabular}       & 2888,48 & 3406,47 & 2704,68 & 3078,63 & 3190,31 & 2963,32 \\
+Banknote                                                       & 31,89   & 12,01   & 17,67   & 22,08   & 21,70   & 18,43   \\
+\begin{tabular}[c]{@{}l@{}}Breast\\ Cancer\end{tabular}        & 9,29    & 7,57    & 6,25    & 8,10    & 8,50    & 5,96    \\
+CarEvaluation                                                  & 50,90   & 63,33   & 50,95   & 53,07   & 69,48   & 45,23   \\
+CNAE9                                                          & 295,55  & 207,57  & 318,40  & 262,22  & 253,48  & 144,87  \\
+CreditApproval                                                 & 14,24   & 12,56   & 13,37   & 13,40   & 15,78   & 10,78   \\
+Digits                                                         & 132,77  & 57,46   & 82,11   & 111,61  & 92,41   & 62,47   \\
+Ecoli                                                          & 5,54    & 6,30    & 4,04    & 5,39    & 6,36    & 3,63    \\
+Parkinsons                                                     & 3,52    & 5,04    & 3,58    & 3,80    & 4,38    & 2,92    \\
+Semeion                                                        & 218,40  & 119,48  & 157,90  & 135,74  & 193,62  & 114,15  \\
+\begin{tabular}[c]{@{}l@{}}Statlog\\ Segmentation\end{tabular} & 156,05  & 55,43   & 87,07   & 159,16  & 91,57   & 66,47   \\
+Wilt                                                           & 70,04   & 75,76   & 68,47   & 67,41   & 78,18   & 60,91   \\
+Zoo                                                            & 3,02    & 3,90    & 2,99    & 3,08    & 4,21    & 1,78    \\ \hline
+\end{tabular}
+\end{table}
 

### 2023-08-22 2023/ITMO paper/elsarticle-template-num.tex
add author information
fix table

--- 2023-08-21 2023/ITMO paper/elsarticle-template-num.tex
@@
 
+\begin{table}[]
+\caption{14 datasets used in the experiments.}
+\label{tab:1}
+\begin{tabular}{lllllll}
+\hline
+Name                                                            & Num   & Attributes & Int & Float & Categorical & Classes \\ \hline
+Balance                                                         & 625   & 4          & 4   & 0     & 0           & 3       \\
+Bank Marketing                                                  & 45211 & 20         & 6   & 4     & 10          & 2       \\
+Banknote                                                        & 1372  & 4          & 0   & 4     & 0           & 2       \\
+Breast Cancer                                                   & 569   & 9          & 9   & 0     & 0           & 2       \\
+CarEvaluation                                                   & 1728  & 6          & 0   & 0     & 6           & 4       \\
+CNAE9                                                           & 1080  & 856        & 856 & 0     & 0           & 9       \\
+Credit Approval                                                 & 690   & 15         & 0   & 6     & 9           & 2       \\
+Digits                                                          & 5620  & 64         & 64  & 0     & 0           & 10      \\
+Ecoli                                                           & 336   & 7          & 0   & 7     & 0           & 8       \\
+Parkinsons                                                      & 197   & 22         & 0   & 22    & 0           & 2       \\
+Semeion                                                         & 1593  & 256        & 256 & 0     & 0           & 10      \\
+\begin{tabular}[c]{@{}l@{}}Statlog \\ Segmentation\end{tabular} & 2310  & 19         & 1   & 18    & 0           & 7       \\
+Wilt                                                            & 4889  & 5          & 0   & 5     & 0           & 2       \\
+Zoo                                                             & 101   & 16         & 16  & 0     & 0           & 7       \\ \hline
+\end{tabular}
+\end{table}
+
 Metrics
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+\begin{table}[]
+\caption{Настраиваемые гиперпараметры для алгоритмов машинного обучения}
+\label{tab:2}
+\begin{tabular}{llll}
+\hline
+Метод                    & Параметры            & Тип параметра & Диапазон параметра                                                     \\ \hline
+\multirow{7}{*}{XGBoost} & n\_estimators        & Int           & {[}10, 200{]}                                                          \\
+                         & max\_depth           & Int           & {[}5, 20{]}                                                            \\
+                         & min\_child\_weight   & Int           & {[}1, 10{]}                                                            \\
+                         & Gamma                & Float         & {[}0.01, 0.6{]}                                                        \\
+                         & Subsample            & Float         & {[}0.05, 0.95{]}                                                       \\
+                         & colsample\_bytree    & Float         & {[}0.05, 0.95{]}                                                       \\
+                         & learning\_rate       & Float         & {[}0.001, 0.1{]}                                                       \\ \hline
+\multirow{3}{*}{\begin{tabular}[c]{@{}l@{}}Support \\ Vector   \\ Classification \\ (SVC)\end{tabular}} &
+  Gamma &
+  Float &
+  {[}$10^{-9}$, $10^{-6}${]} \\
+                         & C                    & Int           & {[}$10^1$,$10^{10}${]}                     \\
+                         & Kernel               & Categorial    & \begin{tabular}[c]{@{}l@{}}\{‘poly’, ‘rbf’,\\ ‘sigmoid’\}\end{tabular} \\ \hline
+\multirow{4}{*}{\begin{tabular}[c]{@{}l@{}}Multi-layer   \\ Perceptron \\ (MLP)\end{tabular}} &
+  Activation &
+  Categorial &
+  \begin{tabular}[c]{@{}l@{}}\{'identity', \\ 'logistic', \\ 'tanh', 'relu'\}\end{tabular} \\
+                         & hidden\_layer\_sizes & Int           & {[}2, 30{]}                                                            \\
+                         & max\_iter            & Int           & {[}100, 700{]}                                                         \\
+                         & Solver               & Categorial    & \begin{tabular}[c]{@{}l@{}}\{'lbfgs', 'sgd', \\ 'adam'\}\end{tabular}  \\ \hline
+\end{tabular}
+\end{table}
 
@@
 Third result
+
+\begin{table}[]
+\caption{Время поиска значений гиперпараметров при однократном запуске алгоритма оптимизации.}
+\label{tab:3}
+\begin{tabular}{lllllll}
+\hline
+Dataset &
+  \begin{tabular}[c]{@{}l@{}}Optuna \\ random\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Optuna \\ tpe\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Optuna \\ cmaes\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Optuna \\ nsgaii\end{tabular} &
+  \begin{tabular}[c]{@{}l@{}}Hyperopt\\ tpe\end{tabular} &
+  iOpt \\ \hline
+Balance                                                         & 10.8   & 9.2    & 9.0    & 10.6   & 10.3   & 7.7    \\
+\begin{tabular}[c]{@{}l@{}}Bank \\ Marketing\end{tabular}       & 2888.5 & 3406.5 & 2704.7 & 3078.6 & 3190.3 & 2963.3 \\
+Banknote                                                        & 31.9   & 12.0   & 17.7   & 22.1   & 21.7   & 18.4   \\
+\begin{tabular}[c]{@{}l@{}}Breast \\ Cancer\end{tabular}        & 9.3    & 7.6    & 6.3    & 8.1    & 8.5    & 6.0    \\
+CarEvaluation                                                   & 50.9   & 63.3   & 51.0   & 53.1   & 69.5   & 45.2   \\
+CNAE9                                                           & 295.6  & 207.6  & 318.4  & 262.2  & 253.5  & 144.9  \\
+\begin{tabular}[c]{@{}l@{}}Credit \\ Approval\end{tabular}      & 14.2   & 12.6   & 13.4   & 13.4   & 15.8   & 10.8   \\
+Digits                                                          & 132.8  & 57.5   & 82.1   & 111.6  & 92.4   & 62.5   \\
+Ecoli                                                           & 5.5    & 6.3    & 4.0    & 5.4    & 6.4    & 3.6    \\
+Parkinsons                                                      & 3.5    & 5.0    & 3.6    & 3.8    & 4.4    & 2.9    \\
+Semeion                                                         & 218.4  & 119.5  & 157.9  & 135.7  & 193.6  & 114.2  \\
+\begin{tabular}[c]{@{}l@{}}Statlog \\ Segmentation\end{tabular} & 156.1  & 55.4   & 87.1   & 159.2  & 91.6   & 66.5   \\
+Wilt                                                            & 70.0   & 75.8   & 68.5   & 67.4   & 78.2   & 60.9   \\
+Zoo                                                             & 3.0    & 3.9    & 3.0    & 3.1    & 4.2    & 1.8    \\ \hline
+\end{tabular}
+\end{table}
 

### 2023-08-22 2023/ITMO paper/elsarticle-template-num.tex
b table

@@
 \author[UNN,ITMO]{Konstantin Barkalov}
+\author[UNN,ITMO]{Denis Karchkov}
 \author[UNN,ITMO]{Evgeny Kozinov}
 \author[UNN,ITMO]{Ilya Lebedev}
-\author[UNN,ITMO]{Denis Karchkov}
 \author[UNN,ITMO]{Denis Rodionov}
+\author[UNN,ITMO]{Marina Usova}
 
@@
   iOpt \\ \hline
-Balance                                                         & 10.8   & 9.2    & 9.0    & 10.6   & 10.3   & 7.7    \\
-\begin{tabular}[c]{@{}l@{}}Bank \\ Marketing\end{tabular}       & 2888.5 & 3406.5 & 2704.7 & 3078.6 & 3190.3 & 2963.3 \\
-Banknote                                                        & 31.9   & 12.0   & 17.7   & 22.1   & 21.7   & 18.4   \\
-\begin{tabular}[c]{@{}l@{}}Breast \\ Cancer\end{tabular}        & 9.3    & 7.6    & 6.3    & 8.1    & 8.5    & 6.0    \\
-CarEvaluation                                                   & 50.9   & 63.3   & 51.0   & 53.1   & 69.5   & 45.2   \\
-CNAE9                                                           & 295.6  & 207.6  & 318.4  & 262.2  & 253.5  & 144.9  \\
-\begin{tabular}[c]{@{}l@{}}Credit \\ Approval\end{tabular}      & 14.2   & 12.6   & 13.4   & 13.4   & 15.8   & 10.8   \\
-Digits                                                          & 132.8  & 57.5   & 82.1   & 111.6  & 92.4   & 62.5   \\
-Ecoli                                                           & 5.5    & 6.3    & 4.0    & 5.4    & 6.4    & 3.6    \\
-Parkinsons                                                      & 3.5    & 5.0    & 3.6    & 3.8    & 4.4    & 2.9    \\
-Semeion                                                         & 218.4  & 119.5  & 157.9  & 135.7  & 193.6  & 114.2  \\
-\begin{tabular}[c]{@{}l@{}}Statlog \\ Segmentation\end{tabular} & 156.1  & 55.4   & 87.1   & 159.2  & 91.6   & 66.5   \\
-Wilt                                                            & 70.0   & 75.8   & 68.5   & 67.4   & 78.2   & 60.9   \\
-Zoo                                                             & 3.0    & 3.9    & 3.0    & 3.1    & 4.2    & 1.8    \\ \hline
+Balance                                                         & 10.8   & 9.2    & 9.0    & 10.6   & 10.3   & \textbf{7.7}    \\
+\begin{tabular}[c]{@{}l@{}}Bank \\ Marketing\end{tabular}       & 2888.5 & 3406.5 & \textbf{2704.7} & 3078.6 & 3190.3 & 2963.3 \\
+Banknote                                                        & 31.9   & \textbf{12.0}   & 17.7   & 22.1   & 21.7   & 18.4   \\
+\begin{tabular}[c]{@{}l@{}}Breast \\ Cancer\end{tabular}        & 9.3    & 7.6    & 6.3    & 8.1    & 8.5    & \textbf{6.0}    \\
+CarEvaluation                                                   & 50.9   & 63.3   & 51.0   & 53.1   & 69.5   & \textbf{45.2}   \\
+CNAE9                                                           & 295.6  & 207.6  & 318.4  & 262.2  & 253.5  & \textbf{144.9}  \\
+\begin{tabular}[c]{@{}l@{}}Credit \\ Approval\end{tabular}      & 14.2   & 12.6   & 13.4   & 13.4   & 15.8   & \textbf{10.8}   \\
+Digits                                                          & 132.8  & \textbf{57.5}   & 82.1   & 111.6  & 92.4   & 62.5   \\
+Ecoli                                                           & 5.5    & 6.3    & 4.0    & 5.4    & 6.4    & \textbf{3.6}    \\
+Parkinsons                                                      & 3.5    & 5.0    & 3.6    & 3.8    & 4.4    & \textbf{2.9}    \\
+Semeion                                                         & 218.4  & 119.5  & 157.9  & 135.7  & 193.6  & \textbf{114.2}  \\
+\begin{tabular}[c]{@{}l@{}}Statlog \\ Segmentation\end{tabular} & 156.1  & \textbf{55.4}   & 87.1   & 159.2  & 91.6   & 66.5   \\
+Wilt                                                            & 70.0   & 75.8   & 68.5   & 67.4   & 78.2   & \textbf{60.9}   \\
+Zoo                                                             & 3.0    & 3.9    & 3.0    & 3.1    & 4.2    & \textbf{1.8}    \\ \hline
 \end{tabular}
@@
 
-Konstantin Barkalov:
-Denis Rodionov:
+\textbf{Konstantin Barkalov:} Supervision, Conceptualization, Writing - original draft.
+\textbf{Denis Karchkov:} Software, Data curation.
+\textbf{Evgeny Kozinov:} Methodology, Validation.
+\textbf{Ilya Lebedev:} Software, Investigation.
+\textbf{Denis Rodionov:} Software, Data curation.
+\textbf{Marina Usova:} Visualization.
 

### 2023-08-24 2023/MMST/samplepaper.tex
add template for paper to MMST

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%
\usepackage{amsmath}

\begin{document}
%
\title{An algorithm for finding the global extremum 
of a partially defined function\thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Marina Usova\inst{1}\orcidID{0000-0002-0722-6884} \and
Konstantin Barkalov\inst{1}\orcidID{0000-0001-5273-2471}}
%
\authorrunning{M. Usova et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603022 Nizhni Novgorod, Russia
\email{\{usova,konstantin.barkalov\}@itmm.unn.ru}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The paper discusses the problem of finding the global minimum of a function that may be partially defined in the search domain. The objective function may be not defined at some search regions due to the nature of the optimized object and the simulation method (for example, numerical instability of the simulation method used). In some cases, these regions are known; but in the most cases information about them is missing. The paper gives a description of the global search algorithm for solving such class of problems. Numerical experiments confirming the efficiency of the proposed algorithm are carried out.

\keywords{Lipschitz global optimization \and multiextremal functions \and hidden constraints \and dimensionality reduction.}
\end{abstract}
%
%
%
\section{Introduction}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\section{Problem Statement}
...
\begin{equation}
\phi^*=\phi(y^* )=\min_{y \in D} \phi(y), D=\left\{ y \in R^N: a_i \leq y_i \leq b_i, a,b \in R^N,1 \leq i \leq N \right\},
\end{equation}
where $y=(y_1,y_2,...,y_N)$ - ..., $D$-dimensional hypercube, $N$ – problem dimension.
... 
$$
D=\left\{ y \in R^N: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\}
$$
...
\begin{equation}
\phi^*=\phi(y(x^* ))=\min_{x \in [0,1]} \left\{ \phi(y(x)) \right\},
\end{equation}
...
\begin{equation}
| \phi (y')-\phi (y'') | \leq L \| y'-y'' \|, y',y' \in D,
\end{equation}
...
\begin{equation}
| \phi (x')-\phi (x'') | \leq K \| x'-x'' \| ^{1/N}, x',x' \in [0,1],
\end{equation}
... $K \leq 4L\sqrt N$ ...
...

\section{Method}

\subsection{Discription of ... }

\subsubsection{Global search algorithm (GSA)} ... $\phi (x)$ ... $x^1 \in (0,1)$ ... $x^{k+1}, k \leq 1$ ... $(k+1)$

\paragraph{Rule 1} Renumber (by the lower indices) the points $x^i, 1 \leq i \leq k$  ...
\begin{equation}
0=x_0 < x_1 < ... < x_i < ... < x_k=1
\end{equation}
... $z_i=\phi (x_i), 1 \leq i \leq k$ ... $v_i=v(x_i)$ ...
\begin{equation}
v_i=v(x_i)=
  \begin{cases}
    -1 & {\quad \text{if } x_i \text{ is boundary point}}\\
    1  & {\quad \text{if } x_i \text{ is internal point}}
  \end{cases}
\end{equation}
... $x_0=0$ and $x_{k+1}=1$ ... $z_0$ and $z_{k+1}$ ...

\paragraph{Rule 2} Compute current lower estimate
\begin{equation}
\mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i}, 2 \leq i \leq k \right\} , \Delta _i= \| x_i -x_{i-1} \| =(x_i-x_{i-1})^{1/N},
\end{equation}
... 
\paragraph{Rule 3} ...
\begin{equation}
z^*=\min \left\{ \phi (x_i): 1\leq i \leq k \right\}.
\end{equation}

\paragraph{Rule 4} For each interval $(x_{i-1},x_i),1 \leq i \leq k+1$ compute the value $R(i)$ called a characteristic of the interval according to the formula
\begin{equation}
R(x_i)=
  \begin{cases}
    \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1}\\
    2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1}\\
    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1}
  \end{cases}
\end{equation}

\paragraph{Rule 5} Select interval $t$ with the largest value of characteristic $R(i):R(t)= \max_{1 \leq i \leq k+1} R(i)$.
\paragraph{Rule 6} ... $(x_{t-1},x_t)$ ...
\begin{equation}
x^{k+1}=\frac {x_t+x_{t-1}}{2}, v(x_{t-1})\neq v(x_t),
\end{equation}
...
\begin{equation}
x^{k+1}= \frac {x_t+x_{t-1}}{2} -  \text{sign} {(z_t-z_{t-1})} \frac {{|z_t-z_{t-1}|}^N}{2r \mu ^N}, v(x_{t-1})=v(x_t).
\end{equation}
...

\subsection{Imputation... }
...
\begin{itemize}
  \item Item 1
  \item Item 2
  \item Item 3
  \item Item 4
  \item Item 5
  \item Item 6
\end{itemize}
...
\begin{equation}
v_i=v(x_i)=
  \begin{cases}
    -1 & {\quad \text{if } x_i \text{ is boundary point}}\\
    1  & {\quad \text{if } x_i \text{ is internal point}}\\
    0  & {\quad \text{if } x_i \text{ is non-computability}}
  \end{cases}
\end{equation}
...

\subsubsection{First...} ...
\begin{equation}
R(i)=\Delta _i,v(x_i)=0 \text{ or } v(x_{i-1})=0
\end{equation}
... 
\begin{figure}
\includegraphics[width=\textwidth]{fig1.pdf}
\caption{Examples ... } \label{fig1}
\end{figure}
\begin{figure}
\includegraphics[width=\textwidth]{fig1.pdf}
\caption{Examples ... } \label{fig2}
\end{figure}

\subsubsection{Second...} ... 
\begin{equation}
R(i)={(1- \frac {1}{r})}^N \Delta _i,
\end{equation}
where $v(x_i)=v(x_{i-1})=0 \text{ or } v(x_i)=0 \text{ and } v(x_{i-1})=-1  \text{ or } v(x_i)=-1 \text{ and } v(x_{i-1})=0$, ...
\begin{equation}
R(i)=2\Delta _i-4 \frac {(z_i-z^*)}{r \mu}, v(x_{i-1})=-1  \text{ or } v(x_{i-1})=0, v(x_i)=1,
\end{equation}
\begin{equation}
R(i)=2\Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, v(x_{i-1})=1, v(x_i)=-1 \text{ or } v(x_i)=0,
\end{equation}
... on ~\ref{fig1} and ~\ref{fig2} ...

\subsubsection{Third...} ...
\begin{figure}
\includegraphics[width=\textwidth]{fig1.pdf}
\caption{Examples ... } \label{fig3}
\end{figure}
...

\subsubsection{Fourth...} ...
\begin{equation}
\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
\end{equation}
where ... 
\begin{figure}
\includegraphics[width=\textwidth]{fig1.pdf}
\caption{Examples ... } \label{fig4}
\end{figure}
\begin{figure}
\includegraphics[width=\textwidth]{fig1.pdf}
\caption{Examples ... } \label{fig5}
\end{figure}

\subsubsection{Fifth...} ...
\paragraph{Step 1} ...
\begin{equation}
\hat{x}',\hat{x}''=\frac {x'+x''}{2}\pm \text{sign}(z''-z')\cdot \frac {1}{2r} \cdot {\left(\frac {z''-z'}{\mu}\right)}^N,
\end{equation}
\paragraph{Step 2} ...
\begin{equation}
\tilde{z}(x)=
  \begin{cases}
    z', & {\quad x \leq \hat{x}',}\\
    \frac {z'+z''}{2}- \frac {r \mu}{2} (\| x - x' \| + \| x - x'' \|), & {\quad \hat{x}' < x < \hat{x}'',}\\
    z'',  & {\quad x \geq \hat{x}''.}
  \end{cases}
\end{equation}
...

\subsubsection{Discriprion of modification} ...
\begin{enumerate} 
  \item ...
  \item ...
  \item ...
\begin{equation}
z_v^*=
  \begin{cases}
    -\varepsilon _r, & {\quad v=0 ,}\\
    \min \left\{ \phi (x_i): v(x_i)=1, 1 \leq i \leq k \right\}, & {\quad v=1 ,}
  \end{cases}
\end{equation}
where $\varepsilon _r$ - ...

  \item ... $(x_{i-1}, x_i), 1 \leq i \leq k+1$ ...
      \begin{enumerate} 
        \item Case 1. ...
\begin{equation}
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z_1^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1}\\
    2 \Delta _i-4 \frac {(z_i-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1}\\
    2 \Delta _i-4 \frac {(z_{i-1}-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1}
  \end{cases}
\end{equation}
        \item Case 2. ...
\begin{equation}
R(i)=
  \begin{cases}
    2 \Delta _i-4 \frac {(z_{i+1}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=0}\\
    2 \Delta _i-4 \frac {(z_{i-2}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=-1}
  \end{cases}
\end{equation}
        \item Case 3. ...
\begin{equation}
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-2})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-2}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=1}\\
     \Delta _i+\frac {{(z_{i+1}-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_{i+1}+z_{i-1}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0}
  \end{cases}
\end{equation}
        \item Case 4. ...
\begin{equation}
R(i)=\Delta _i+\frac {{(\tilde{z}_i-\tilde{z}_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {\tilde{z}_i+\tilde{z}_{i-1}-2z_1^*}{r \mu}, v(x_{i-1})=v(x_i)=0
\end{equation}
        \item Case 5. ...
\begin{equation}
R(i)=\Delta _i \cdot {\left( 1-\frac {1}{r} \right)}^N+z_0^*.
\end{equation}
    \end{enumerate} 
  \item ...
  \item ...
\end{enumerate}

\subsection{Results... }
...  ~\ref{tab1} ...
\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|l|}
\hline
 &  GKLS classical & GKLS round district & GKLS random district \\
\hline
Average number of iterations & 781,66	& 854,91 & 839,03 \\
Average number of non-computability & - & 185,79 & 42,31 \\
Number of solved problems & 100 / 100 & 100 / 100 & 100 / 100\\
\hline
\end{tabular}
\end{table}

\begin{figure}
\includegraphics[width=\textwidth]{fig1.pdf}
\caption{...} \label{fig6}
\end{figure}
...
\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|l|l|}
\hline
 &  & Average number of iterations & Average number of non-computability & Solved problems \\
\hline
$N=3$ & classical & 12251,11 & - & 100 / 100 \\
$N=3$ & round district & 12201,78 & 4763,57 & 100 / 100 \\
$N=3$ & random district & 12210,71 & 63,39 & 100 / 100 \\
$N=4$ & classical & 25144,46 & - & 100 / 100 \\
$N=4$ & round district & 29840,32 & 15004,37  & 100 / 100 \\
$N=4$ & random district & 25287,49 & 26,43 & 100 / 100 \\
$N=5$ & classical & 56725,95 & - & 100 / 100 \\
$N=5$ & round district & 47840,09 & 24109,19 & 100 / 100\\
$N=5$ & random district & 57874,76 & 82,57 & 100 / 100\\
\hline
\end{tabular}
\end{table}
...
\subsection{Conclusion}
...

 (see Fig.~\ref{fig1}).

For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.

\subsubsection{Acknowledgements} Please place your acknowledgments at
the end of the paper, preceded by an unnumbered run-in heading (i.e.
3rd-level heading).

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}

### 2023-08-24 2023/ITMO paper/elsarticle-template-num.tex
add some references

@@
 \usepackage{multirow}
+
+\usepackage{hyperref}
 
@@
 В настоящее время сложные научно и/или промышленные задачи чаще всего решаются численными методами.
-В большинстве таких методов, например, в методах машинного обучения, имеются гиперпараметры, от выбора конкретных значений которых может зависеть эффективность (в заданной метрике) полученного решения при заданных тем или иным способом ограничениях на время расчетов \cite{Hutter2019}.
+В большинстве таких методов, например, в методах машинного обучения, имеются гиперпараметры, от выбора конкретных значений которых может зависеть эффективность (в заданной метрике) полученного решения при заданных тем или иным способом ограничениях на время расчетов \cite{Hutter2019,nikitin2021}.
 Тот факт, что во многих случаях разница в эффективности получаемого решения при разных значениях гиперпараметров может быть достаточно значительной, привел к возникновению the hyperparameter optimization (HPO) problem, а также исследованиям подходов к ее решению, реализуемых в фреймворках автоматического подбора гиперпараметров \cite{Tune, optuna, hyperopt,Sherpa}.
@@
 
-В данной статье мы представляем фреймворк методов интеллектуальной оптимизации iOpt, ориентированный на решение задач выбора оптимальных значений гиперпараметров различных алгоритмов (включая методы машинного обучения и эвристической оптимизации).
+В данной статье мы представляем фреймворк методов интеллектуальной оптимизации iOpt\footnote{\url{https://github.com/aimclub/iOpt/}}, ориентированный на решение задач выбора оптимальных значений гиперпараметров различных алгоритмов (включая методы машинного обучения и эвристической оптимизации).
 Статья построена следующим образом.
@@
 \textbf{Denis Karchkov:} Software, Data curation.
-\textbf{Evgeny Kozinov:} Methodology, Validation.
-\textbf{Ilya Lebedev:} Software, Investigation.
-\textbf{Denis Rodionov:} Software, Data curation.
+\textbf{Evgeny Kozinov:} Methodology, Validation, Writing - original draft.
+\textbf{Ilya Lebedev:} Software, Investigation, Writing - original draft.
+\textbf{Denis Rodionov:} Software, Investigation.
 \textbf{Marina Usova:} Visualization.
@@
 
+\section*{Code availability}
+%проверить перевод
+Software that implements all of the described methods is available in the open repository \url{https://github.com/aimclub/iOpt/}.
+
 \section*{Acknowledgements}
-
+%уточнить ссылку
 This research was supported by the research center ``Strong Artificial Intelligence in Industry'' of ITMO University.

### 2023-08-27 2023/MMST/samplepaper.tex
add english text and pictures, add biblio

@@
 The paper discusses the problem of finding the global minimum of a function that may be partially defined in the search domain.
-The objective function may be not defined at some search regions due to the nature of the optimized object and the simulation method (for example, numerical instability of the simulation method used).
-In some cases, these regions are known; but in the most cases information about them is missing.
-The paper gives a description of the global search algorithm for solving such class of problems.
-Numerical experiments confirming the efficiency of the proposed algorithm are carried out.
+The objective function may not be defined within some search regions due to the nature of the optimized object and of the simulation method (for example, numerical instability of the simulation method used).
+In some cases, these regions are known; but in most cases, the information about them is missing.
+The paper gives a description of a global search algorithm for solving such class of problems.
+Numerical experiments confirming the efficiency of the proposed algorithm were carried out.
 
@@
 \section{Introduction}
-Please note that the first paragraph of a section or subsection is
-not indented. The first paragraph that follows a table, figure,
-equation etc. does not need an indent, either.
+Global optimization [...rus...] problems of finding the global (absolute) minima of multidimensional multiextremal functions.
+The possibility of building a reliable estimate of the global extremum in such problems is based principally on the availability of the \textit{a priori} known information on the problem to be solved allowing relating the possible values of the objective function to the known values in the points of search trials performed.
 
-Subsequent paragraphs, however, are indented.
+Often, such information is represented in the form of suggestion that the objective function $\phi$ satisfies the Lipschitz condition with a priori unknown constant $L$.
+
+This suggestion can be interpreted (in the relation to the applied problems) as the reflection of a limited power generating the changes in the simulated system.
+At that, the objective function defined in the form of a "\textit{black box}"\ type functions as a rule is a non differentiable one, and each computation of its value in some point within a feasible domain may require considerable computational resources.
+
+A number of efficient deterministic algorithms for solving the Lipschitz global optimization problems have been developed ~\cite{Jones2009,Birect2020,Sergeyev2017}.
+The comparisons conducted have shown the deterministic algorithms to overcome (in various criteria) the widespread nature-inspired algorithms ~\cite{Liberti2005,Sergeyev2018,Sovrasov2019} .
+
+The present work continues the development of one of the efficient deterministic algorithms for solving the Lipschitz global optimization problems, namely the of the information-statistical algorithm of global search ~\cite{indexMethod,strongin1978,Strongin2000}.
+This algorithm suggests the solving of a multidimensional problem by means of its reduction to an equivalent one-dimensional optimization problem using the space-filling curves (Peano curves) ~\cite{Sergeyev2013}.
+
+Note that the use of complex mathematical models is natural in the optimization of complex real world objects that, as a consequence, increases the computation costs for searching the optimum Essentially.
+In last decades, the experts in the fields of optimization and parallel computations have proposed many methods for reducing the computation costs an speed-up of the algorithms related to solving the optimization problems appearing as well as to the numerical analysis of the original models [...ref???...].
+
+However, a principally new problem in applications, namely the numerical instability of the models investigated in some (unknown a priori) regions in the ranges of the parameters’ variations became relevant recently.
+In such regions, one cannot perform the numerical modeling and compute the objective function values correctly.
+This phenomenon can be interpreted either as the presence of some hidden (unknown a priori) constraints In the problem or as a partial (limited) definition of the objective function in the search domain.
+In such a statement, the optimization problem becomes more complex essentially since the region of feasible parameters’ combinations is not defined in advance.
+
+This paper presents the results in a new research direction related to the use of imputation\footnote{The term «imputation» in the paper is used following the established terminology in the field of machine learning related to the restoring of the missing data.} (restoring) of the objective function values in the course of the global search algorithm operation when solving the problems with partially defined objective functions.
+The approach proposed is based on an adoption of the computation rules of the global search algorithm.
+The efficiency of the approach proposed has been demonstrated numerically.
 
 \section{Problem Statement}
-...
-\begin{equation}
+In the general form, the global optimization problem can be formulated as follows:
+\begin{equation}\label{eq1} 
 \phi^*=\phi(y^* )=\min_{y \in D} \phi(y), D=\left\{ y \in R^N: a_i \leq y_i \leq b_i, a,b \in R^N,1 \leq i \leq N \right\},
 \end{equation}
-where $y=(y_1,y_2,...,y_N)$ - ..., $D$-dimensional hypercube, $N$ – problem dimension.
-... 
+where $y=(y_1,y_2,...,y_N)$ - is the vector of the varied parameters, $D$ - is an $N$-dimensional hypercube, and $N$ –  is the dimension of the problem being solved.
+
+Using the Peano curves mapping the interval $[0,1]$ onto the $N$-dimensional unitary hypercube unambiguously
 $$
-D=\left\{ y \in R^N: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\}
+D=\left\{ y \in R^N: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\},
 $$
-...
-\begin{equation}
+we can reduce the initial problem (\ref{eq1}) to a one-dimensional problem
+\begin{equation}\label{eq2} 
 \phi^*=\phi(y(x^* ))=\min_{x \in [0,1]} \left\{ \phi(y(x)) \right\},
 \end{equation}
-...
-\begin{equation}
+that allows applying efficient one-dimensional optimization algorithms for its solving.
+
+We make the following assumptions on the objective function.
+
+First, the objective function may be multiextremal, nondifferentiable, and, moreo-ver, defined in the form of a "black box"\ (i.
+e.
+in the form of some subroutine, which entry the argument is supplied to, and the output is respective function value).
+
+Second, each computing of the function in some point of the feasible region may require considerable computation resources.
+
+Third, the objective function satisfies the Lipschitz condition
+\begin{equation}\label{eq3} 
 | \phi (y')-\phi (y'') | \leq L \| y'-y'' \|, y',y' \in D,
 \end{equation}
-...
-\begin{equation}
+where $L$ is the Lipschitz constant.
+The dimensionality reduction scheme using Peano curves is known to juxtapose the multidimensional problem with a Lipschitz objective function (\ref{eq1}) to problem (\ref{eq2}) with the one-dimensional objective function satisfying H{\"o}lder condition
+\begin{equation}\label{eq4} 
 | \phi (x')-\phi (x'') | \leq K \| x'-x'' \| ^{1/N}, x',x' \in [0,1],
 \end{equation}
-... $K \leq 4L\sqrt N$ ...
-...
+where $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the relation $K \leq 4L\sqrt N$ \cite{Strongin2000}.
 
-\section{Method}
+Various versions of algorithms for solving the problems of such a class and respective theories of convergence are presented in \cite{Gourdin1996,Lera2015,strongin1978}.
 
-\subsection{Discription of ... }
+Fourth, the objective function may be undefined in some regions of search domain $D$ (in particular cases, in one or in several points of this one).
 
-\subsubsection{Global search algorithm (GSA)} ... $\phi (x)$ ... $x^1 \in (0,1)$ ... $x^{k+1}, k \leq 1$ ... $(k+1)$
+From our experience in solving the applied optimization problems [...refs??...], we will assume the total volume of the non-computability regions to be a small fraction of the whole volume of the search domain $D$ (about 10-20\%)\footnote{When testing the algorithm, we have considered also more complex problem statements, which the volume of the non-computability region achieved 50-80\% in.
+The problems of this kind were solved by the algorithm proposed successfully as well but a large number of trials was required to solve the problem because the method attempted to explore the whole non-computability region.
+In this connection, in the rules of constructing the problems the experiments for the dimensionality $N=5$ were modified so that the non-computability region were located in the corners of the search domain only.}.
 
-\paragraph{Rule 1} Renumber (by the lower indices) the points $x^i, 1 \leq i \leq k$  ...
-\begin{equation}
+The last assumption makes the application of the classic information-statistical \textit{global search algorithm} \cite{indexMethod,strongin1978} or other Lipschitz optimization methods impossible and requires developing a modification for solving the problems with the objective functions, which are not defined everywhere.
+
+\section{Solving Method}
+
+\subsection{Description of Classic Global Search Algorithm}
+Within the framework of the approach proposed, the information-statistical theory of global search is used for solving the problems (\ref{eq2}).
+The main idea of the algorithm is the use of the accumulated information to determine the next interval, which the absolute minimum is most probable to find within.
+The next trial point in this interval corresponds to the mathematical expectation of the minimum location.
+
+Let us give more detailed description of the algorithm.
+
+\subsubsection{Global search algorithm (GSA).} At every iteration of the global search a \textit{search trial} is executed.
+We will call a trial a computing the optimized function value  $\phi (x)$ from (\ref{eq2}).
+The first trial will be performed in the middle internal point $x^1 \in (0,1)$.
+The choice of the point $x^{k+1}, k \leq 1$ of the next $(k+1)^\text{th}$ trial is performed on the base of the following rules.
+
+\paragraph{Rule 1.} Renumber (by the lower indices) the points $x^i, 1 \leq i \leq k$  of the previous trials in the ascending order by their coordinates i. e.
+\begin{equation}\label{eq5} 
 0=x_0 < x_1 < ... < x_i < ... < x_k=1
 \end{equation}
-... $z_i=\phi (x_i), 1 \leq i \leq k$ ... $v_i=v(x_i)$ ...
-\begin{equation}
+and juxtapose these ones to the values $z_i=\phi (x_i), 1 \leq i \leq k$ computed in these points and to the indices $v_i=v(x_i)$ defined by the rule
+\begin{equation}\label{eq6} 
 v_i=v(x_i)=
   \begin{cases}
-    -1 & {\quad \text{if } x_i \text{ is boundary point}}\\
-    1  & {\quad \text{if } x_i \text{ is internal point}}
+    -1 & {\quad \text{if } x_i \text{ is a boundary point,}}\\
+    1  & {\quad \text{if } x_i \text{ is an internal point,}}
   \end{cases}
 \end{equation}
-... $x_0=0$ and $x_{k+1}=1$ ... $z_0$ and $z_{k+1}$ ...
+the points $x_0=0$ and $x_{k+1}=1$ are introduced additionally (the values $z_0$ and $z_{k+1}$ are undefined) for convenience of the following notations.
 
-\paragraph{Rule 2} Compute current lower estimate
-\begin{equation}
+\paragraph{Rule 2.} Compute current lower estimate
+\begin{equation}\label{eq7} 
 \mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i}, 2 \leq i \leq k \right\} , \Delta _i= \| x_i -x_{i-1} \| =(x_i-x_{i-1})^{1/N},
 \end{equation}
-... 
-\paragraph{Rule 3} ...
-\begin{equation}
+of the unknown H{\"o}lder constant $K$ from (\ref{eq4}) for the reduced function $\phi(x)$ where $r, r>1$ is a parameter of the algorithm; if $\mu=0$ accept $\mu=1$.
+\paragraph{Rule 3.} Determine the best current value of the objective function
+\begin{equation}\label{eq8} 
 z^*=\min \left\{ \phi (x_i): 1\leq i \leq k \right\}.
@@
 
-\paragraph{Rule 4} For each interval $(x_{i-1},x_i),1 \leq i \leq k+1$ compute the value $R(i)$ called a characteristic of the interval according to the formula
-\begin{equation}
+\paragraph{Rule 4.} For each interval $(x_{i-1},x_i),1 \leq i \leq k+1$ compute the value $R(i)$ called a \textit{characteristic} of the interval according to the formula
+\begin{equation}\label{eq9} 
 R(x_i)=
@@
 
-\paragraph{Rule 5} Select interval $t$ with the largest value of characteristic $R(i):R(t)= \max_{1 \leq i \leq k+1} R(i)$.
-\paragraph{Rule 6} ... $(x_{t-1},x_t)$ ...
-\begin{equation}
+\paragraph{Rule 5.} Select the interval $t$ with the largest value of characteristic $R(i):R(t)= \max_{1 \leq i \leq k+1} R(i)$.
+
+\paragraph{Rule 6.} Perform the next trial in the middle point of the interval $(x_{t-1},x_t)$ if the indices of its end points are not the same i. e.
+\begin{equation}\label{eq10} 
 x^{k+1}=\frac {x_t+x_{t-1}}{2}, v(x_{t-1})\neq v(x_t),
 \end{equation}
-...
-\begin{equation}
+Otherwise, perform the next trial in the point
+\begin{equation}\label{eq11} 
 x^{k+1}= \frac {x_t+x_{t-1}}{2} -  \text{sign} {(z_t-z_{t-1})} \frac {{|z_t-z_{t-1}|}^N}{2r \mu ^N}, v(x_{t-1})=v(x_t).
 \end{equation}
-...
 
-\subsection{Imputation... }
-...
+\textit{Stop criterion} of the algorithm, according to which the execution of the algorithm is terminated for current global optimization problem (\ref{eq2}) is the achievement of a predefined accuracy of the problem solution i.
+e.
+if the length of the interval $(x_{t-1},x_t)$ containing current point $x^{k+1}$ is less or equal to the predefined accuracy $\Delta _t \leq \varepsilon$ where $t$ is from Rule 5 and $\varepsilon >0$ is the predefined accuracy.
+
+\subsection{ Approaches to the Reconstruction of the Objective Function Values }
+The problem of not everywhere defined objective function described above is a partial analog of the long-standing missing data problem in Data Science.
+The origin of this problem may be, for example, the technical problems or assembling the data set from several sources with different sets of parameters.
+
+It is worth noting that some machine learning algorithms can take into account and even restore the missing values in the data.
+For example, LightGBM has a regime of ignoring the missing values while XGBoost restores the data by means of the de-crease in the loss function when learning.
+Nevertheless, the majority of models can-not process the missing values and require the valid data without NaN or «missing» values.
+It makes this problem to be serious enough.
+
+The experts in Data Science have proposed various approaches to the imputation (restoring) of the missing data:
 \begin{itemize}
-  \item Item 1
-  \item Item 2
-  \item Item 3
-  \item Item 4
-  \item Item 5
-  \item Item 6
+  \item Deleting of all cases, which have missing values of some variables (the last resort, the algorithm loses the access to the useful information contained in the on-missing observations’ values);
+  \item Replacing any missing value with the average of the available values of the variable (doesn’t work with the qualitative variables, doesn’t account for the correlation of the data);
+  \item Replacing by the most frequent value or by a constant (differs from the use of the median only by the ability to work with the qualitative parameters);
+  \item Replacing the data using the method of $k$ nearest neighbors or kNN (unlike the former, accounts for the correlation between the parameters, sensitive to the fadeouts, computation costly);
+  \item Multiple imputation by chained equations (MICE) – imputation of each value is performed not one but many times (allows understanding how reliable or unreliable is the value suggested);
+  \item Imputation of the data using a deep learning (restores the missing values by means of training a neural network on the points, which all parameters are available for, computation costly for big data).
 \end{itemize}
-...
-\begin{equation}
+
+The absence of an universal algorithm and the necessity to find the most suitable methods for each particular problem are the common drawbacks of the approaches listed above.
+
+For the imputation of the objective function values, one can consider, in general, a similar set of approaches, however, taking into account the Lipschitz condition and using the information-statistical theory.
+It allows selecting a universal algorithm fro this problem class.
+Next, the points, which the non-computability was found in are interpreted as the ones having the zero index, and (\ref{eq6}) is replaced by an extended formula
+\begin{equation}\label{eq12} 
 v_i=v(x_i)=
@@
 \end{equation}
-...
 
-\subsubsection{First...} ...
-\begin{equation}
-R(i)=\Delta _i,v(x_i)=0 \text{ or } v(x_{i-1})=0
+Let us describe various possible approaches to the implementation of the modification.
+
+\subsubsection{The first approach.}To ignore the absence of the objective function value in the point.
+The characteristic of an interval with even a single non-computable point is computed from the length of the interval
+\begin{equation}\label{eq13} 
+R(i)=\Delta _i, \text{ if } v(x_i)=0 \text{ or } v(x_{i-1})=0.
 \end{equation}
-... 
+
 \begin{figure}
-\includegraphics[width=\textwidth]{fig1.pdf}
-\caption{Examples ... } \label{fig1}
+\includegraphics[width=\textwidth]{fig1.png}
+\caption{Examples of the distributions of the search trials in the non-computability regions in the first approach to the imputation of the values} \label{fig1}
+\end{figure}
+
+A large number of excess trials after getting into the non-computability region due to a large value of the characteristic obtained according to (\ref{eq13}) is an essential draw-back of this approach.
+
+\subsubsection{The second approach.} To ignore the absence of the objective function values in the point.
+The characteristic of the interval with two non-computable point or with one non-computable point and one boundary point is computed from the interval multiplied by a reducing constant
+\begin{equation}\label{eq14} 
+R(i)={(1- \frac {1}{r})}^N \Delta _i,
+\end{equation}
+if $v(x_i)=v(x_{i-1})=0$ or $v(x_i)=0$ and $v(x_{i-1})=-1$ or $v(x_i)=-1$ and $v(x_{i-1})=0$.
+The characteristic of the interval with one non-computable point and one computable point is calculated according to the rules of operation with the boundary interval
+\begin{equation}\label{eq15} 
+R(i)=2\Delta _i-4 \frac {(z_i-z^*)}{r \mu},\text{ if } v(x_{i-1})=-1  \text{ or } v(x_{i-1})=0, v(x_i)=1,
+\end{equation}
+\begin{equation}\label{eq16} 
+R(i)=2\Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu},\text{ if } v(x_{i-1})=1, v(x_i)=-1 \text{ or } v(x_i)=0.
+\end{equation}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{fig2.png}
+\caption{Examples of the distributions of the search trials in the non-computability regions in the second approach to the imputation of the values} \label{fig2}
+\end{figure}
+
+This approach resolves partly the problems of the first approach outlined above.
+However, it doesn’t solve the problem of excess trials and requires a considerable increasing of the parameter $r$ for more complex test problem.
+The examples for the one-dimensional problems using the first two approaches are presented in ~\ref{fig1} and ~\ref{fig2} where the trials falling into the region with the undefined values are marked by the black points and the computable points of the search domain – by the green ones.
+
+\subsubsection{The third approach.} To restore the objective function values in the points of the intervals with two non-computable points as the averages of the neighbor points and to apply corresponding computation rule from formula (\ref{eq9}) assuming the points to be computable ones.
+If the neighbor points are non-computable as well, to use formula (\ref{eq14}).
+In the case of an interval with one non-computable point, to apply formulae (\ref{eq15}) and (\ref{eq16}).
+\begin{figure}
+\includegraphics[width=\textwidth]{fig3.png}
+\caption{Imputation of the non-computable values by the averages of the neighbor points. Approach \#3} \label{fig3}
+\end{figure}
+
+Essential drawbacks of this approach are not utilizing the information on the coordinates aw well as seldom executing the computation rule added because of searching the substituting values among the nearest neighbors only.
+
+\subsubsection{The fourth approach.} To restore the objective function values in the points for the intervals with two non-computable points as the averages of the nearest computable points taking into account the location coordinates i.
+e.
+\begin{equation}\label{eq17} 
+\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
+\end{equation}
+where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval.
+The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located near the boundary point only.
+\begin{figure}
+\includegraphics[width=\textwidth]{fig4.png}
+\caption{Imputation of non-computable values by averages of the neighboring points. Approach \#4} \label{fig4}
 \end{figure}
 \begin{figure}
-\includegraphics[width=\textwidth]{fig1.pdf}
-\caption{Examples ... } \label{fig2}
+\includegraphics[width=\textwidth]{fig5.png}
+\caption{Imputation of non-computable values by averages of the minorant and the majorant. Approach \#5} \label{fig5}
 \end{figure}
 
-\subsubsection{Second...} ... 
-\begin{equation}
-R(i)={(1- \frac {1}{r})}^N \Delta _i,
-\end{equation}
-where $v(x_i)=v(x_{i-1})=0 \text{ or } v(x_i)=0 \text{ and } v(x_{i-1})=-1  \text{ or } v(x_i)=-1 \text{ and } v(x_{i-1})=0$, ...
-\begin{equation}
-R(i)=2\Delta _i-4 \frac {(z_i-z^*)}{r \mu}, v(x_{i-1})=-1  \text{ or } v(x_{i-1})=0, v(x_i)=1,
-\end{equation}
-\begin{equation}
-R(i)=2\Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, v(x_{i-1})=1, v(x_i)=-1 \text{ or } v(x_i)=0,
-\end{equation}
-... on ~\ref{fig1} and ~\ref{fig2} ...
-
-\subsubsection{Third...} ...
-\begin{figure}
-\includegraphics[width=\textwidth]{fig1.pdf}
-\caption{Examples ... } \label{fig3}
-\end{figure}
-...
-
-\subsubsection{Fourth...} ...
-\begin{equation}
-\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
-\end{equation}
-where ... 
-\begin{figure}
-\includegraphics[width=\textwidth]{fig1.pdf}
-\caption{Examples ... } \label{fig4}
-\end{figure}
-\begin{figure}
-\includegraphics[width=\textwidth]{fig1.pdf}
-\caption{Examples ... } \label{fig5}
-\end{figure}
-
-\subsubsection{Fifth...} ...
-\paragraph{Step 1} ...
-\begin{equation}
+\subsubsection{The fifth approach.} To utilize the search information and the suggestions on the function form most completely, one can restore the objective function values in the points taking onto account the H{\"o}lder condition.
+\paragraph{Step 1.} Find the end points of the crossing of majorant and minorant
+\begin{equation}\label{eq18} 
 \hat{x}',\hat{x}''=\frac {x'+x''}{2}\pm \text{sign}(z''-z')\cdot \frac {1}{2r} \cdot {\left(\frac {z''-z'}{\mu}\right)}^N,
 \end{equation}
-\paragraph{Step 2} ...
-\begin{equation}
+\paragraph{Step 2.} Restore the objective function values according to the rule
+\begin{equation}\label{eq19} 
 \tilde{z}(x)=
@@
 \end{equation}
-...
+This rule can be interpreted as an average value of the minorant and the majorant for a function satisfying the H{\"o}lder condition.
+Let us give the description of the modification of the global search algorithm in the case of not everywhere computable function based on this approach.
 
-\subsubsection{Discriprion of modification} ...
+\subsubsection{Description of the modification.} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions:
 \begin{enumerate} 
-  \item ...
-  \item ...
-  \item ...
-\begin{equation}
+  \item The index $v_i=v(x_i)$ is determined by the formula (\ref{eq12}).
+  \item When computing current lower estimate of the H{\"o}lder constant $K$ in formula (\ref{eq7}), use the intervals with two computable points only.
+  \item The best current objective function value is determined for different values of the index
+\begin{equation}\label{eq20} 
 z_v^*=
@@
 \end{equation}
-where $\varepsilon _r$ - ...
+where $\varepsilon _r$ is a parameter of the method.
 
-  \item ... $(x_{i-1}, x_i), 1 \leq i \leq k+1$ ...
-      \begin{enumerate} 
-        \item Case 1. ...
-\begin{equation}
+  \item For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$, the computing of the characteristic $R(i)$ is performed according to the following computing formulae.
+      \begin{itemize} 
+        \item \textit{Case 1.} For the intervals with \textbf{two computable points} or with \textbf{one computable point and one boundary point}, apply the classic rules of global search algorithm
+\begin{equation}\label{eq21} 
 R(i)=
@@
 \end{equation}
-        \item Case 2. ...
-\begin{equation}
+        \item \textit{Case 2.}  For the intervals with \textbf{one boundary point and one non-computable point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
+\begin{equation}\label{eq22} 
 R(i)=
@@
 \end{equation}
-        \item Case 3. ...
-\begin{equation}
+        \item \textit{Case 3.} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor com-putable point and apply the formula
+\begin{equation}\label{eq23} 
 R(i)=
@@
 \end{equation}
-        \item Case 4. ...
-\begin{equation}
+        \item \textit{Case 4.}.
+For the intervals with \textbf{two non-computable points}, perform the imputation of the values for the non-computable points based on the rule (\ref{eq19}) and apply the formula
+\begin{equation}\label{eq24} 
 R(i)=\Delta _i+\frac {{(\tilde{z}_i-\tilde{z}_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {\tilde{z}_i+\tilde{z}_{i-1}-2z_1^*}{r \mu}, v(x_{i-1})=v(x_i)=0
 \end{equation}
-        \item Case 5. ...
-\begin{equation}
+        \item \textit{Case 5.} In the case when any imputation is impossible, use the formula
+\begin{equation}\label{eq25} 
 R(i)=\Delta _i \cdot {\left( 1-\frac {1}{r} \right)}^N+z_0^*.
 \end{equation}
-    \end{enumerate} 
-  \item ...
-  \item ...
+    \end{itemize} 
+  \item If several intervals have equal characteristics, prefer the computable interval with the lowest number.
+  \item If the best interval appears to be the one with two non-computable points, perform the next trial in the middle point of the interval.
 \end{enumerate}
 
-\subsection{Results... }
-...  ~\ref{tab1} ...
+\subsection{Results of Numerical Experiments}
+The numerical experiments were carried out using a computer with Intel® Core™ i7-10750H CPU @ 2.60GHz 2.59 GHz using the developed software package implementing the modification described in Section 3.
+To generate a series of problems, GKLS generator described in \cite{Gaviano2003} was used.
+It allows generating the multiextremal oprimization problems with the properties known in advance.
+These problems were appended with the non-computable regions of various kinds.
+
+Each series of experiments consisted of 100 GKLS problems with everywhere computable objective function (the classic GKLS problems), 100 problems with the non-computabilities at the boundaries of the search domains, and 100 problems with non-computability regions defined randomly.
+
+In the first series of experiments, the correctness of the modification operation was checked and the features of executing the trials in the non-computability regions were revealed.
+The problems of the dimensionality $N=2$ were used in the experiment.
+The reliability parameter was set to $r=5.5$, the precision $\varepsilon=0.01$.
+The results of experiments are presented in Tab.~\ref{tab1}.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{fig6.png}
+\caption{Examples of problems \#\# 19, 25, and 35 from the first series of experiments} \label{fig6}
+\end{figure}
+
+Fig.~\ref{fig6} presents the examples of the modification operation with several problems of the series.
+The trials inside the computable regions are denoted by the blue points in the figures, the global minima of the problems – by the red points, and the points of trials performed inside the non-computability regions – by black.
+Also, the non-computability regions are filled with grey for clarity.
+
+Since the modification works with the classic GKLS problem series according to the base global search algorithm, one can compare it with the classic GSA.
+Thus, the general behavior of GSA was preserved for the modification, and the same regions of trials’ densification near the local minima were observed that can be explained by the realization of the classical computation rules (\ref{eq21}) in most cases.
+One can see in Tab.~\ref{tab1} that the number of trials executed inside the non-computable regions is small.
+Total number of trials for the classic problems and for the problems with non-computable regions are comparable, the method doesn’t perform excess iterations in the non-promising regions due to the computing formulae utilizing the information on the nearest computable points.
+
 \begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|l|}
+\caption{Comparison of the efficiencies of the algorithm modification with the classic version of the global search algorithm on GKLS problems}\label{tab1}
+\begin{tabular}{|l|c|c|c|}
 \hline
- &  GKLS classical & GKLS round district & GKLS random district \\
+ &  classical  &  round district  &  random district  \\
 \hline
@@
 
-\begin{figure}
-\includegraphics[width=\textwidth]{fig1.pdf}
-\caption{...} \label{fig6}
-\end{figure}
-...
+The use of rule (\ref{eq25}) forms a grid of trials inside the non-computability regions that is visible clearly for Problem \#35.
+At that, the trial points are denser at the boundary of the non-computable region close to potential minimum when attempting to approach this one from different sides due to the computational rules (\ref{eq22})-(\ref{eq24}) that was observed in all problems presented as the examples.
+
+100\% of the problems were solved.
+However, for the series with the boundary non-computability region, the value of parameter $r=6.0$ was required to solve one problem, for the series with the random regions – soving of two problems with given value of parameter $r$.
+
+In the second series of experiments, the problem series of various dimensionalities were solved.
+In the experiments with the problems of the dimensionality $N=4$, the reliability parameter was set to $r=4.5$, the precision $\varepsilon=0.02$.
+The global minimum was considered to be found if the algorithm generates a trial point inside the $\varepsilon$-nearness of the global minimum.
+The results of the experiments are presented in Tab.~\ref{tab2}.
+
 \begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|l|l|}
+\caption{Problem series of various dimensionalities}\label{tab2}
+\begin{tabular}{|l|l|c|c|c|}
 \hline
- &  & Average number of iterations & Average number of non-computability & Solved problems \\
+ N &  Task type & Average num. of iters & Average num. of non-comput. & Solved problems \\
 \hline
-$N=3$ & classical & 12251,11 & - & 100 / 100 \\
-$N=3$ & round district & 12201,78 & 4763,57 & 100 / 100 \\
-$N=3$ & random district & 12210,71 & 63,39 & 100 / 100 \\
-$N=4$ & classical & 25144,46 & - & 100 / 100 \\
-$N=4$ & round district & 29840,32 & 15004,37  & 100 / 100 \\
-$N=4$ & random district & 25287,49 & 26,43 & 100 / 100 \\
-$N=5$ & classical & 56725,95 & - & 100 / 100 \\
-$N=5$ & round district & 47840,09 & 24109,19 & 100 / 100\\
-$N=5$ & random district & 57874,76 & 82,57 & 100 / 100\\
+$3$ & classical & 12251,11 & - & 100 / 100 \\
+$ $ & round & 12201,78 & 4763,57 & 100 / 100 \\
+$ $ & random & 12210,71 & 63,39 & 100 / 100 \\
+\hline
+$4$ & classical & 25144,46 & - & 100 / 100 \\
+$ $ & round & 29840,32 & 15004,37  & 100 / 100 \\
+$ $ & random & 25287,49 & 26,43 & 100 / 100 \\
+\hline
+$5$ & classical & 56725,95 & - & 100 / 100 \\
+$ $ & round & 47840,09 & 24109,19 & 100 / 100\\
+$ $ & random & 57874,76 & 82,57 & 100 / 100\\
 \hline
@@
 \end{table}
-...
-\subsection{Conclusion}
-...
 
- (see Fig.~\ref{fig1}).
+According to the results of the experiments, the method resolves successfully the occasional falls into the non-computability regions scattered inside the search domain randomly whereas in the case of large boundary regions it tries the execute a complete investigation without executing too many excess trials.
 
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
+\subsection{Conclusions}
+In the present work, the problem of finding the global minimum of a computation-costly "black box"\ function was considered.
+As compared to the traditional formulation used in the global optimization, the problem considered has an important differ-ence: the objective function may be undefined in some regions of the search domain (in particular, in one point or in several ones).
+The appearing of such a problem statement is caused by a frequent problem of numerical instability of the models investigated in the applied problems.
 
-\subsubsection{Acknowledgements} Please place your acknowledgments at
-the end of the paper, preceded by an unnumbered run-in heading (i.e.
-3rd-level heading).
+The main problem is the impossibility to identify the non-computability regions in advance in most cases.
+In fact, the region of feasible combinations of parameters is undefined a priori.
+This phenomenon can be interpreted as the presence of additional hidden constraints in the problem.
+
+The problem investigated in the present work echoes to problem of missing data existing in Data Science for a long time.
+As a solution, the experts in the field have proposed various approaches to the imputation (restoring) of the missing values.
+These ideas have found their reflection in the development of the new method.
+
+The applicability of already existing approaches appended with various variants to account for the Lipschitz condition was investigated.
+As a result, a universal efficient algorithm for solving the problems of this class has been developed.
+The algorithm constructed is a development of one of efficient deterministic method of solving the Lipschitz global optimization problems, namely information-statistical global search algorithm.
+Within the framework of this modification, the computation rules of the algorithm were adopted with introduction of the data imputation algorithm.
+
+[....rus??? ...] $N = 2,3,4,5$ ...
+
+%\subsubsection{Acknowledgements} Please place your acknowledgments at
+%the end of the paper, preceded by an unnumbered run-in heading (i.e.
+%3rd-level heading).
 
@@
 %
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
+\bibliographystyle{splncs04}
+\bibliography{bibliography}
 
-\bibitem{ref_lncs1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
 \end{document}

### 2023-08-27 2023/MMST/samplepaper.tex
fix titles and itemize

@@
 \end{equation}
-where $y=(y_1,y_2,...,y_N)$ - is the vector of the varied parameters, $D$ - is an $N$-dimensional hypercube, and $N$ –  is the dimension of the problem being solved.
+where $y=(y_1,y_2,...,y_N)$ is the vector of the varied parameters, $D$ is an $N$-dimensional hypercube, and $N$ is the dimension of the problem being solved.
 
@@
 
-From our experience in solving the applied optimization problems [...refs??...], we will assume the total volume of the non-computability regions to be a small fraction of the whole volume of the search domain $D$ (about 10-20\%)\footnote{When testing the algorithm, we have considered also more complex problem statements, which the volume of the non-computability region achieved 50-80\% in.
+From our experience in solving the applied optimization problems [refs??], we will assume the total volume of the non-computability regions to be a small fraction of the whole volume of the search domain $D$ (about 10-20\%)\footnote{When testing the algorithm, we have considered also more complex problem statements, which the volume of the non-computability region achieved 50-80\% in.
 The problems of this kind were solved by the algorithm proposed successfully as well but a large number of trials was required to solve the problem because the method attempted to explore the whole non-computability region.
@@
 
-\subsubsection{Global search algorithm (GSA).} At every iteration of the global search a \textit{search trial} is executed.
+\subsubsection{Global search algorithm (GSA)} At every iteration of the global search a \textit{search trial} is executed.
 We will call a trial a computing the optimized function value  $\phi (x)$ from (\ref{eq2}).
@@
 
-\paragraph{Rule 1.} Renumber (by the lower indices) the points $x^i, 1 \leq i \leq k$  of the previous trials in the ascending order by their coordinates i. e.
+\paragraph{Rule 1} Renumber (by the lower indices) the points $x^i, 1 \leq i \leq k$  of the previous trials in the ascending order by their coordinates i. e.
 \begin{equation}\label{eq5} 
@@
 
-\paragraph{Rule 2.} Compute current lower estimate
+\paragraph{Rule 2} Compute current lower estimate
 \begin{equation}\label{eq7} 
@@
 of the unknown H{\"o}lder constant $K$ from (\ref{eq4}) for the reduced function $\phi(x)$ where $r, r>1$ is a parameter of the algorithm; if $\mu=0$ accept $\mu=1$.
-\paragraph{Rule 3.} Determine the best current value of the objective function
+\paragraph{Rule 3} Determine the best current value of the objective function
 \begin{equation}\label{eq8} 
@@
 
-\paragraph{Rule 4.} For each interval $(x_{i-1},x_i),1 \leq i \leq k+1$ compute the value $R(i)$ called a \textit{characteristic} of the interval according to the formula
+\paragraph{Rule 4} For each interval $(x_{i-1},x_i),1 \leq i \leq k+1$ compute the value $R(i)$ called a \textit{characteristic} of the interval according to the formula
 \begin{equation}\label{eq9} 
@@
 
-\paragraph{Rule 5.} Select the interval $t$ with the largest value of characteristic $R(i):R(t)= \max_{1 \leq i \leq k+1} R(i)$.
+\paragraph{Rule 5} Select the interval $t$ with the largest value of characteristic $R(i):R(t)= \max_{1 \leq i \leq k+1} R(i)$.
 
-\paragraph{Rule 6.} Perform the next trial in the middle point of the interval $(x_{t-1},x_t)$ if the indices of its end points are not the same i. e.
+\paragraph{Rule 6} Perform the next trial in the middle point of the interval $(x_{t-1},x_t)$ if the indices of its end points are not the same i. e.
 \begin{equation}\label{eq10} 
@@
 
-\textit{Stop criterion} of the algorithm, according to which the execution of the algorithm is terminated for current global optimization problem (\ref{eq2}) is the achievement of a predefined accuracy of the problem solution i.
+\paragraph{Stop criterion} of the algorithm, according to which the execution of the algorithm is terminated for current global optimization problem (\ref{eq2}) is the achievement of a predefined accuracy of the problem solution i.
 e.
@@
 \subsubsection{The fifth approach.} To utilize the search information and the suggestions on the function form most completely, one can restore the objective function values in the points taking onto account the H{\"o}lder condition.
-\paragraph{Step 1.} Find the end points of the crossing of majorant and minorant
+\paragraph{Step 1} Find the end points of the crossing of majorant and minorant
 \begin{equation}\label{eq18} 
@@
 \end{equation}
-\paragraph{Step 2.} Restore the objective function values according to the rule
+\paragraph{Step 2} Restore the objective function values according to the rule
 \begin{equation}\label{eq19} 
@@
 
-\subsubsection{Description of the modification.} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions:
+\subsubsection{Description of the modification} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions:
 \begin{enumerate} 
@@
   \item For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$, the computing of the characteristic $R(i)$ is performed according to the following computing formulae.
-      \begin{itemize} 
-        \item \textit{Case 1.} For the intervals with \textbf{two computable points} or with \textbf{one computable point and one boundary point}, apply the classic rules of global search algorithm
+
+        \paragraph{Case 1} For the intervals with \textbf{two computable points} or with \textbf{one computable point and one boundary point}, apply the classic rules of global search algorithm
 \begin{equation}\label{eq21} 
@@
 \end{equation}
-        \item \textit{Case 2.}  For the intervals with \textbf{one boundary point and one non-computable point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
+        \paragraph{Case 2}   For the intervals with \textbf{one boundary point and one non-computable point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
 \begin{equation}\label{eq22} 
@@
 \end{equation}
-        \item \textit{Case 3.} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor com-putable point and apply the formula
+        \paragraph{Case 3} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor com-putable point and apply the formula
 \begin{equation}\label{eq23} 
@@
 \end{equation}
-        \item \textit{Case 4.}.
-For the intervals with \textbf{two non-computable points}, perform the imputation of the values for the non-computable points based on the rule (\ref{eq19}) and apply the formula
+        \paragraph{Case 4} For the intervals with \textbf{two non-computable points}, perform the imputation of the values for the non-computable points based on the rule (\ref{eq19}) and apply the formula
 \begin{equation}\label{eq24} 
@@
 \end{equation}
-        \item \textit{Case 5.} In the case when any imputation is impossible, use the formula
+        \paragraph{Case 5} In the case when any imputation is impossible, use the formula
 \begin{equation}\label{eq25} 
@@
 \end{equation}
-    \end{itemize} 
   \item If several intervals have equal characteristics, prefer the computable interval with the lowest number.
@@
 
-[....rus??? ...] $N = 2,3,4,5$ ...
+[....rus...] $N = 2,3,4,5$ ...
 

### 2023-08-27 2023/vestnik.susu/example.tex
Исправлена часть замечаний

@@
 \usepackage{upgreek}
+\usepackage{multirow}
 
@@
 
-% ЗАПОЛНЯЕТСЯ АВТОРАМИ: Релевантные коды УДК статьи
-% (см. http://teacode.com/online/udc/)
-\classify{УДК  519.853.4}
+\classify{} % Код MSC не нужен
 
@@
 
-\title{Об использовании деревьев решений для выявления областей притяжения локальных минимумов в параллельном алгоритме глобальной оптимизации%
-	\footnote{Статья рекомендована к публикации программным комитетом Международной научной конференции <<Параллельные вычислительные технологии (ПаВТ) 2023>>}%
+\title{ON USING THE DECISION TREES TO IDENTIFY THE LOCAL EXTREMA IN PARALLEL GLOBAL OPTIMIZATION ALGORITHMS
+\footnote{The paper is recommended for publication by the Program Committee of the International Scientific Conference “Parallel Computational Technologies (PCT) 2023”}
 }
 
-\authors{К.А.~Баркалов$^{1}$, И.Г.~Лебедев$^{1}$, Д.И.~Силенко$^{1}$}
+\authorsen{Konstantin Barkalov, Ilya Lebedev,  Dmitry Silenko}
 
-\organizations{$^{1}$Нижегородский государственный университет им. Н.И. Лобачевского\\
-	(603022, Н.Новгород, проспект Гагарина, ~23)}
+\organizations{Lobachevsky State University of Nizhny Novgorod\\
+	(23 Prospekt Gagarina, Nizhny Novgorod, 603022 Russia)}
 
@@
 % ЗАПОЛНЯЕТСЯ АВТОРАМИ: Дата поступления статьи в редакцию 
-\newcommand{\rdate}{28.07.2023}
+\newcommand{\rdate}{31.07.2023}
 \emails{\emailsinsert}
-\redactiondate{\rdate}
+\redactiondateen{\rdate}
 
@@
 \setcounter{footnote}{0}
-
-% Оформление аннотации
-\begin{abstract}%
-В работе рассматривается решение многомерных задач многоэкстремальной оптимизации с использованием деревьев решений для выявления областей притяжения локальных минимумов.
-Целевая функция представлена как «черный ящик»; для нее лишь предполагается, что она удовлетворяет условию Липшица с априори неизвестной константой.
-Для решения задачи многоэкстремальной оптимизации применятся алгоритм глобального поиска.
-Предложена модификация алгоритма, в которой определяются окрестности локальных минимумов целевой функции на основе анализа накопленной поисковой информации.
-Проведение такого анализа с использованием методов машинного обучения позволяет принять решение о запуске локального метода, что может ускорить сходимость алгоритма.
-Данный подход был подтвержден результатами численных экспериментов, демонстрирующих ускорение при решении набора тестовых задач.
-
-\keywords{Глобальная оптимизация, Многоэкстремальные функции, Параллельные вычисления, Машинное обучение, Дерево решений}
-\end{abstract}
-
-%Образец цитирования статьи
-\forcitation{Баркалов К.А., Лебедев  И.Г., Силенко  Д.И. Об использовании деревьев решений для выявления областей притяжения локальных минимумов в параллельном алгоритме глобальной оптимизации}
-
-%Верхние колонтитулы
-\markboth{К.А.~Баркалов, И.Г.~Лебедев, Д.И.~Силенко}{Об использовании деревьев решений в параллельном алгоритме глобальной оптимизации \ldots}
-
-\vspace{1em}
-{\it Работа выполнена при поддержке Российского научного фонда, проект № 21-11-00204 .}
-
-\insertbiblio
-
-% Сведения об авторах (Ф.И.О.(полностью), степень, звание, кафедра (отдел), вуз, e-mail)
-Баркалов Константин Александрович, д.т.н., доцент, Кафедра математического обеспечения и суперкомпьютерных технологий, Нижегородский государственный университет им.
-Н.И.
-Лобачевского (Н.Новгород, Российская Федерация)
-
-Лебедев Илья Генадьевич, заведующий лабораторией суперкомпьютерных технологий и высокопроизводительных вычислений,  Нижегородский государственный университет им.
-Н.И.
-Лобачевского (Н.Новгород, Российская Федерация)
-
-Силенко Дмитрий Игоревич, студент, Нижегородский государственный университет им. Н.И. Лобачевского (Н.Новгород, Российская Федерация)
-
-%\newpage
-
-{\phantom{eee} \hrule \vskip 4 pt}
-\classify{} % Код MSC не нужен
-%\classify{MSC 00000} % Коды MSC см. http://www.ams.org/msc/
-
-\title{On using the decision trees to identify the local extrema in parallel global optimization algorithms}
-
-\authorsen{Konstantin Barkalov$^{1}$, Ilya Lebedev$^{1}$,  Dmitry Silenko$^{1}$}
-
-\organizations{$^{1}$Lobachevsky State University of Nizhny Novgorod\\
-	(23 Prospekt Gagarina, Nizhny Novgorod, 603022 Russia)}
-
-\emails{\emailsinsert}
-\redactiondateen{\rdate}
 
@@
 
-\keywordsen{Global optimization, Multiextremal functions, Parallel computing, Machine learning, Decision tree}
+	\keywordsen{global optimization, multiextremal functions, parallel computing, machine learning, decision tree}
 \end{abstract}
@@
 % Заголовки разделов формируются при помощи команд  \section{}, \subsection{}, \subsubsection{}
-\section{Introduction}
+\section*{Introduction}
 
@@
 		\hline
-		N & Problem class & DIRECT & DIRECTl & GSA  \\ \hline
-		4 & Simple & $>47282 (4)$ & 18983 & 11953  \\ \hline
-		~ & Hard & $>95708 (7)$ & 68754 & 25263  \\ \hline
-		5 & Simple & $>16057 (1)$ & 16758 & 15920  \\ \hline
-		~ & Hard & $>217215 (16)$ & $>269064 (4)$ & $>148342 (4)$  \\ \hline
+		$N$                & Problem class & DIRECT         & DIRECTl       & GSA           \\ \hline
+		\multirow{2}{*}{4} & Simple        & $>47282 (4)$   & 18983         & 11953         \\ \cline{2-5} 
+		& Hard          & $>95708 (7)$   & 68754         & 25263         \\ \hline
+		\multirow{2}{*}{5} & Simple        & $>16057 (1)$   & 16758         & 15920         \\ \cline{2-5} 
+		& Hard          & $>217215 (16)$ & $>269064 (4)$ & $>148342 (4)$ \\ \hline
 	\end{tabular}
@@
 	\centering
-	\begin{tabular}{|c|c|c|c|c|c|}
+	\begin{tabular}{|l|l|ll|ll|}
 		\hline
-
-		N & Problem class & \multicolumn{2}{c|}{Averaged number of iterations} & \multicolumn{2}{c|}{Speedup} \\ \hline
-		& ~ & AGSA & Decision tree & AGSA & Decision tree \\ \hline
-		& Simple & 3823.1   & 485.6  & 7.4   & 12.7 \\ \hline
-		2  & Hard & 787.6    & 304.2  & 13.8  & 14.9 \\ \hline
-		& Simple & 5700.5   & 720.1  & 5.6   & 14.2  \\ \hline
-		3  & Hard & 2411.9   & 634.3  & 5.0   & 5.9  \\ \hline
-		& Simple & 31101.5  & 1037.4 & 5.1   & 7.1  \\ \hline
-		4  & Hard & 10418.1  & 816.2  & 6.9   & 5.0  \\ \hline
-		& Simple & 163313.5 & 1204.2 & 4.8   & 4.1 \\ \hline
-		5  & Hard & 27524.5  & 999.4  & 4.6   & 6.3  \\ \hline
+		\multirow{2}{*}{$N$} & \multirow{2}{*}{Problem class} & \multicolumn{2}{l|}{Averaged number   of iterations} & \multicolumn{2}{l|}{Speedup}              \\ \cline{3-6} 
+		&                                & \multicolumn{1}{l|}{AGSA}         & Decision tree    & \multicolumn{1}{l|}{AGSA} & Decision tree \\ \hline
+		\multirow{2}{*}{2}   & Simple                         & \multicolumn{1}{l|}{3823.1}       & 485.6            & \multicolumn{1}{l|}{7.4}  & 12.7          \\ \cline{2-6} 
+		& Hard                           & \multicolumn{1}{l|}{787.6}        & 304.2            & \multicolumn{1}{l|}{13.8} & 14.9          \\ \hline
+		\multirow{2}{*}{3}   & Simple                         & \multicolumn{1}{l|}{5700.5}       & 720.1            & \multicolumn{1}{l|}{5.6}  & 14.2          \\ \cline{2-6} 
+		& Hard                           & \multicolumn{1}{l|}{2411.9}       & 634.3            & \multicolumn{1}{l|}{5}    & 5.9           \\ \hline
+		\multirow{2}{*}{4}   & Simple                         & \multicolumn{1}{l|}{31101.5}      & 1037.4           & \multicolumn{1}{l|}{5.1}  & 7.1           \\ \cline{2-6} 
+		& Hard                           & \multicolumn{1}{l|}{10418.1}      & 816.2            & \multicolumn{1}{l|}{6.9}  & 5             \\ \hline
+		\multirow{2}{*}{5}   & Simple                         & \multicolumn{1}{l|}{163313.5}     & 1204.2           & \multicolumn{1}{l|}{4.8}  & 4.1           \\ \cline{2-6} 
+		& Hard                           & \multicolumn{1}{l|}{27524.5}      & 999.4            & \multicolumn{1}{l|}{4.6}  & 6.3           \\ \hline
 	\end{tabular}
@@
 
-\section{Conclusion}
+\section*{Conclusion}
 
@@
 
+%\vspace{1em}
+%{\it Работа выполнена при поддержке Российского научного фонда, проект № 21-11-00204 .}
+
 \insertbiblioen
+\selectlanguage{russian}
+
+{\phantom{eee} \hrule \vskip 4 pt}
+
+%\classify{MSC 00000} % Коды MSC см. http://www.ams.org/msc/
+
+% ЗАПОЛНЯЕТСЯ АВТОРАМИ: Релевантные коды УДК статьи
+% (см. http://teacode.com/online/udc/)
+\classify{УДК  519.853.4}
+
+\title{ОБ ИСПОЛЬЗОВАНИИ ДЕРЕВЬЕВ РЕШЕНИЙ ДЛЯ ВЫЯВЛЕНИЯ ОБЛАСТЕЙ ПРИТЯЖЕНИЯ ЛОКАЛЬНЫХ МИНИМУМОВ В ПАРАЛЛЕЛЬНОМ АЛГОРИТМЕ ГЛОБАЛЬНОЙ ОПТИМИЗАЦИИ%
+	\footnote{Статья рекомендована к публикации программным комитетом Международной научной конференции <<Параллельные вычислительные технологии (ПаВТ) 2023>>}%
+}
+
+\authors{К.А.~Баркалов, И.Г.~Лебедев, Д.И.~Силенко}
+
+\organizations{Нижегородский государственный университет им. Н.И. Лобачевского\\
+	(603022, Н.Новгород, проспект Гагарина, ~23)}
+
+\emails{\emailsinsert}
+\redactiondate{\rdate}
+
+% Оформление аннотации
+\begin{abstract}%
+	В работе рассматривается решение многомерных задач многоэкстремальной оптимизации с использованием деревьев решений для выявления областей притяжения локальных минимумов.
+Целевая функция представлена как «черный ящик»; для нее лишь предполагается, что она удовлетворяет условию Липшица с априори неизвестной константой.
+Для решения задачи многоэкстремальной оптимизации применятся алгоритм глобального поиска.
+Предложена модификация алгоритма, в которой определяются окрестности локальных минимумов целевой функции на основе анализа накопленной поисковой информации.
+Проведение такого анализа с использованием методов машинного обучения позволяет принять решение о запуске локального метода, что может ускорить сходимость алгоритма.
+Данный подход был подтвержден результатами численных экспериментов, демонстрирующих ускорение при решении набора тестовых задач.
+
+	\keywords{глобальная оптимизация, многоэкстремальные функции, параллельные вычисления, машинное обучение, дерево решений}
+\end{abstract}
+
+%Образец цитирования статьи
+\forcitation{Баркалов К.А., Лебедев  И.Г., Силенко  Д.И. Об использовании деревьев решений для выявления областей притяжения локальных минимумов в параллельном алгоритме глобальной оптимизации}
+
+%Верхние колонтитулы
+\markboth{К.А.~Баркалов, И.Г.~Лебедев, Д.И.~Силенко}{Об использовании деревьев решений в параллельном алгоритме глобальной оптимизации \ldots}
+
+\insertbiblio
+
+% Сведения об авторах (Ф.И.О.(полностью), степень, звание, кафедра (отдел), вуз, e-mail)
+Баркалов Константин Александрович, д.т.н., доцент, Кафедра математического обеспечения и суперкомпьютерных технологий, Нижегородский государственный университет им.
+Н.И.
+Лобачевского (Н.Новгород, Российская Федерация)
+
+Лебедев Илья Генадьевич, заведующий лабораторией суперкомпьютерных технологий и высокопроизводительных вычислений,  Нижегородский государственный университет им.
+Н.И.
+Лобачевского (Н.Новгород, Российская Федерация)
+
+Силенко Дмитрий Игоревич, магистр, Нижегородский государственный университет им. Н.И. Лобачевского (Н.Новгород, Российская Федерация)
+
+%\newpage
 

### 2023-08-27 2023/MMST/samplepaper.tex
add two references

@@
 \title{An algorithm for finding the global extremum 
-of a partially defined function\thanks{Supported by organization x.}}
+of a partially defined function\thanks{This work was supported by the Russian Science Foundation, project No. 21-11-00204.}}
 %
@@
 \section{Introduction}
-Global optimization [...rus...] problems of finding the global (absolute) minima of multidimensional multiextremal functions.
-The possibility of building a reliable estimate of the global extremum in such problems is based principally on the availability of the \textit{a priori} known information on the problem to be solved allowing relating the possible values of the objective function to the known values in the points of search trials performed.
+Global optimization deals with the development of methods for solving problems of finding the global (absolute) minima of multidimensional multiextremal functions.
+The possibility of constructing a reliable estimate of the global extremum in such problems is principally based on the availability of the \textit{a priori} known information on the problem to be solved.
+This information makes it possible to relate the unknown values of the objective function to the known values at the points of the performed search trials.
 
@@
 Note that the use of complex mathematical models is natural in the optimization of complex real world objects that, as a consequence, increases the computation costs for searching the optimum Essentially.
-In last decades, the experts in the fields of optimization and parallel computations have proposed many methods for reducing the computation costs an speed-up of the algorithms related to solving the optimization problems appearing as well as to the numerical analysis of the original models [...ref???...].
+In last decades, the experts in the fields of optimization and parallel computations have proposed many methods for reducing the computation costs an speed-up of the algorithms related to solving the appearing optimization problems \cite{Kvasov2013}  as well as to the numerical analysis of the original models \cite{Duwe2020}.
 

### 2023-08-27 2023/MMST/samplepaper.tex
one more reference

@@
 Note that the use of complex mathematical models is natural in the optimization of complex real world objects that, as a consequence, increases the computation costs for searching the optimum Essentially.
-In last decades, the experts in the fields of optimization and parallel computations have proposed many methods for reducing the computation costs an speed-up of the algorithms related to solving the appearing optimization problems \cite{Kvasov2013}  as well as to the numerical analysis of the original models \cite{Duwe2020}.
+In last decades, the experts in the fields of optimization and parallel computations have proposed many methods for reducing the computation costs an speed-up of the algorithms related to solving the appearing optimization problems \cite{Kvasov2013}  as well as to the numerical analysis of the original models \cite{Dongarra2022,Duwe2020}.
 

### 2023-08-28 2023/MMST/samplepaper.tex
add last paragraph, fix typos and fix to 15 pages

@@
 
-First, the objective function may be multiextremal, nondifferentiable, and, moreo-ver, defined in the form of a "black box"\ (i.
+First, the objective function may be multiextremal, nondifferentiable, and, moreover, defined in the form of a "black box"\ (i.
 e.
@@
 \end{equation}
-where $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the relation $K \leq 4L\sqrt N$ \cite{Strongin2000}.
+where $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the rule $K \leq 4L\sqrt N$ \cite{Strongin2000}.
 
@@
 It is worth noting that some machine learning algorithms can take into account and even restore the missing values in the data.
-For example, LightGBM has a regime of ignoring the missing values while XGBoost restores the data by means of the de-crease in the loss function when learning.
-Nevertheless, the majority of models can-not process the missing values and require the valid data without NaN or «missing» values.
+For example, LightGBM has a regime of ignoring the missing values while XGBoost restores the data by means of the decrease in the loss function when learning.
+Nevertheless, the majority of models cannot process the missing values and require the valid data without NaN or «missing» values.
 It makes this problem to be serious enough.
@@
 \end{equation}
-        \paragraph{Case 3} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor com-putable point and apply the formula
+        \paragraph{Case 3} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor computable point and apply the formula
 \begin{equation}\label{eq23} 
@@
 In the present work, the problem of finding the global minimum of a computation-costly "black box"\ function was considered.
-As compared to the traditional formulation used in the global optimization, the problem considered has an important differ-ence: the objective function may be undefined in some regions of the search domain (in particular, in one point or in several ones).
+As compared to the traditional formulation used in the global optimization, the problem considered has an important difference: the objective function may be undefined in some regions of the search domain (in particular, in one point or in several ones).
 The appearing of such a problem statement is caused by a frequent problem of numerical instability of the models investigated in the applied problems.
@@
 
-[....rus...] $N = 2,3,4,5$ ...
+The numerical experiments were carried out using series of multiextremal GKLS problems of the dimensionalities $N = 2,3,4,5$.
+The results have demonstrated the reliability and efficiency of the approach proposed.
+The modified algorithm handles successfully the problems, which the non-computability regions are scattered randomly within the search domain.
+In the case of large boundary non-computability regions, the method solves the problems efficiently as well performing a complete investigation of the non-computability regions also.
 

### 2023-08-28 2023/MMST/samplepaper.tex
one more reference

--- 2023-08-27 2023/MMST/samplepaper.tex
@@
 
-From our experience in solving the applied optimization problems [refs??], we will assume the total volume of the non-computability regions to be a small fraction of the whole volume of the search domain $D$ (about 10-20\%)\footnote{When testing the algorithm, we have considered also more complex problem statements, which the volume of the non-computability region achieved 50-80\% in.
+From our experience in solving the applied optimization problems \cite{Barkalov2022,Gubaydullin2022}, we will assume the total volume of the non-computability regions to be a small fraction of the whole volume of the search domain $D$ (about 10-20\%)\footnote{When testing the algorithm, we have considered also more complex problem statements, which the volume of the non-computability region achieved 50-80\% in.
 The problems of this kind were solved by the algorithm proposed successfully as well but a large number of trials was required to solve the problem because the method attempted to explore the whole non-computability region.

### 2023-08-28 2023/MMST/samplepaper.tex
minor corrections

--- 2023-08-28 2023/MMST/samplepaper.tex
@@
 
+\usepackage{hyperref}
+
 \begin{document}
@@
 %
-\authorrunning{M. Usova et al.}
+\authorrunning{M. Usova, K. Barkalov}
 % First names are abbreviated in the running head.
@@
 
-Often, such information is represented in the form of suggestion that the objective function $\phi$ satisfies the Lipschitz condition with a priori unknown constant $L$.
+Often such information is represented in the form of suggestion that the objective function $\phi$ satisfies the Lipschitz condition with a priori unknown constant $L$.
 
 This suggestion can be interpreted (in the relation to the applied problems) as the reflection of a limited power generating the changes in the simulated system.
-At that, the objective function defined in the form of a "\textit{black box}"\ type functions as a rule is a non differentiable one, and each computation of its value in some point within a feasible domain may require considerable computational resources.
+At that, the objective function defined in the form of a ``\textit{black box}'' type functions as a rule is a non differentiable one, and each computation of its value in some point within a feasible domain may require considerable computational resources.
 
@@
 
-The present work continues the development of one of the efficient deterministic algorithms for solving the Lipschitz global optimization problems, namely the of the information-statistical algorithm of global search ~\cite{indexMethod,strongin1978,Strongin2000}.
-This algorithm suggests the solving of a multidimensional problem by means of its reduction to an equivalent one-dimensional optimization problem using the space-filling curves (Peano curves) ~\cite{Sergeyev2013}.
+The present work continues the development of one of the efficient deterministic algorithms for solving the Lipschitz global optimization problems, namely the of the information-statistical global search algorithm ~\cite{indexMethod,strongin1978,Strongin2000}.
+This algorithm suggests the solving of a multidimensional problem by means of its reduction to an equivalent one-dimensional optimization problem using the Peano-Hilbert space-filling curves \cite{Sergeyev2013}.
 
-Note that the use of complex mathematical models is natural in the optimization of complex real world objects that, as a consequence, increases the computation costs for searching the optimum Essentially.
+Note that the use of complex mathematical models is natural in the optimization of complex real world objects that, as a consequence, essentially increases the computation costs for searching the optimum.
 In last decades, the experts in the fields of optimization and parallel computations have proposed many methods for reducing the computation costs an speed-up of the algorithms related to solving the appearing optimization problems \cite{Kvasov2013}  as well as to the numerical analysis of the original models \cite{Dongarra2022,Duwe2020}.
 
-However, a principally new problem in applications, namely the numerical instability of the models investigated in some (unknown a priori) regions in the ranges of the parameters’ variations became relevant recently.
+However, a principally new problem in applications, namely the numerical instability of the models investigated in some (unknown a priori) regions in the ranges of the parameters’ variations recently became relevant.
 In such regions, one cannot perform the numerical modeling and compute the objective function values correctly.
-This phenomenon can be interpreted either as the presence of some hidden (unknown a priori) constraints In the problem or as a partial (limited) definition of the objective function in the search domain.
-In such a statement, the optimization problem becomes more complex essentially since the region of feasible parameters’ combinations is not defined in advance.
+This phenomenon can be interpreted either as the presence of some hidden (unknown a priori) constraints or as a partial computability of the objective function in the search domain.
+In such a statement, the optimization problem becomes more complex since the region of feasible parameters’ combinations is not defined in advance.
 
-This paper presents the results in a new research direction related to the use of imputation\footnote{The term «imputation» in the paper is used following the established terminology in the field of machine learning related to the restoring of the missing data.} (restoring) of the objective function values in the course of the global search algorithm operation when solving the problems with partially defined objective functions.
+This paper presents the results in a new research direction related to the use of imputation\footnote{The term ``imputation'' in the paper is used following the established terminology in the field of machine learning related to the restoring of the missing data.} (restoring) of the objective function values in the course of the global search algorithm operation when solving the problems with partially defined objective functions.
 The approach proposed is based on an adoption of the computation rules of the global search algorithm.
@@
 \section{Problem Statement}
+
 In the general form, the global optimization problem can be formulated as follows:
@@
 
-Using the Peano curves mapping the interval $[0,1]$ onto the $N$-dimensional unitary hypercube unambiguously
+Using the Peano-Hilbert curves unambiguously mapping the interval $[0,1]$ onto the $N$-dimensional unit hypercube 
 $$
@@
 
-First, the objective function may be multiextremal, nondifferentiable, and, moreover, defined in the form of a "black box"\ (i.
-e.
+First, the objective function may be multiextremal, nondifferentiable, and, moreover, defined in the form of a ``black box'' (i.e.
 in the form of some subroutine, which entry the argument is supplied to, and the output is respective function value).
-
-Second, each computing of the function in some point of the feasible region may require considerable computation resources.
+Second, each computing of the function at some point of the feasible region may require considerable computation resources.
 
@@
 \end{equation}
-where $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the rule $K \leq 4L\sqrt N$ \cite{Strongin2000}.
+where $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the rule $K \leq 2L\sqrt {N+3}$ \cite{Strongin2000}.
 
-Various versions of algorithms for solving the problems of such a class and respective theories of convergence are presented in \cite{Gourdin1996,Lera2015,strongin1978}.
+Various versions of algorithms for solving the problems of such a class and respective theories of convergence are presented in \cite{Sergeyev2013,Strongin2000}.
 
-Fourth, the objective function may be undefined in some regions of search domain $D$ (in particular cases, in one or in several points of this one).
+Fourth, the objective function may be undefined in some regions of search domain $D$ (in particular cases, at one or several points).
 
-From our experience in solving the applied optimization problems [refs??], we will assume the total volume of the non-computability regions to be a small fraction of the whole volume of the search domain $D$ (about 10-20\%)\footnote{When testing the algorithm, we have considered also more complex problem statements, which the volume of the non-computability region achieved 50-80\% in.
-The problems of this kind were solved by the algorithm proposed successfully as well but a large number of trials was required to solve the problem because the method attempted to explore the whole non-computability region.
-In this connection, in the rules of constructing the problems the experiments for the dimensionality $N=5$ were modified so that the non-computability region were located in the corners of the search domain only.}.
+From our experience in solving the applied optimization problems \cite{Barkalov2022,Gubaydullin2022}, we will assume the total volume of the non-computability regions to be a small fraction of the whole volume of the search domain $D$ (about 10-20\%)\footnote{When testing the algorithm, we have considered also more complex problems, in which the volume of the non-computability region achieved 50-80\%.
+The problems of this kind were successfully solved by the proposed algorithm as well but a large number of trials was required to do it since the method attempted to explore the whole non-computability region.
+In this connection, in the rules of constructing the test problems for the dimensionality $N=5$ were modified so that the non-computability region were located olny in the corners of the search domain.}.
 
-The last assumption makes the application of the classic information-statistical \textit{global search algorithm} \cite{indexMethod,strongin1978} or other Lipschitz optimization methods impossible and requires developing a modification for solving the problems with the objective functions, which are not defined everywhere.
+The last assumption makes the application of the classic information-statistical \textit{global search algorithm} \cite{Strongin2000} or other Lipschitz optimization methods \cite{Sergeyev2017} impossible and requires developing a modification for solving the problems with the objective functions, which are not defined everywhere.
 
@@
 
-\subsection{Description of Classic Global Search Algorithm}
-Within the framework of the approach proposed, the information-statistical theory of global search is used for solving the problems (\ref{eq2}).
-The main idea of the algorithm is the use of the accumulated information to determine the next interval, which the absolute minimum is most probable to find within.
+\subsection{Description of the Global Search Algorithm}
+
+Within the framework of the proposed approach, the information-statistical theory of global search is used for solving the problems (\ref{eq2}).
+The main idea of the algorithm is the use of the accumulated information to determine the next interval, within which it is most probable to find the absolute minimum.
 The next trial point in this interval corresponds to the mathematical expectation of the minimum location.
@@
 
-\subsubsection{Global search algorithm (GSA)} At every iteration of the global search a \textit{search trial} is executed.
-We will call a trial a computing the optimized function value  $\phi (x)$ from (\ref{eq2}).
-The first trial will be performed in the middle internal point $x^1 \in (0,1)$.
+\subsubsection{Global search algorithm (GSA)} At every iteration of the global search a \textit{trial} is executed.
+We will refer to computing the objective function value  $\phi (x)$ from (\ref{eq2}) as a trial.
+The first trial will be executed in the middle point $x^1 \in (0,1)$.
 The choice of the point $x^{k+1}, k \leq 1$ of the next $(k+1)^\text{th}$ trial is performed on the base of the following rules.
 
-\paragraph{Rule 1} Renumber (by the lower indices) the points $x^i, 1 \leq i \leq k$  of the previous trials in the ascending order by their coordinates i. e.
+\paragraph{Rule 1.} Renumber (by the lower indices) the points $x^i, 1 \leq i \leq k$  of the previous trials in the ascending order by their coordinates, i.e.
 \begin{equation}\label{eq5} 
@@
 \end{equation}
-and juxtapose these ones to the values $z_i=\phi (x_i), 1 \leq i \leq k$ computed in these points and to the indices $v_i=v(x_i)$ defined by the rule
+and juxtapose them with the values $z_i=\phi (x_i), 1 \leq i \leq k$, computed at these points and to the indices $v_i=v(x_i)$ defined by the rule
 \begin{equation}\label{eq6} 
@@
     -1 & {\quad \text{if } x_i \text{ is a boundary point,}}\\
-    1  & {\quad \text{if } x_i \text{ is an internal point,}}
+    1  & {\quad \text{if } x_i \text{ is an internal point.}}
   \end{cases}
 \end{equation}
-the points $x_0=0$ and $x_{k+1}=1$ are introduced additionally (the values $z_0$ and $z_{k+1}$ are undefined) for convenience of the following notations.
+The points $x_0=0$ and $x_{k+1}=1$ are introduced additionally (the values $z_0$ and $z_{k+1}$ are undefined) for convenience of the following notations.
 
-\paragraph{Rule 2} Compute current lower estimate
+\paragraph{Rule 2.} Compute the estimate
 \begin{equation}\label{eq7} 
-\mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i}, 2 \leq i \leq k \right\} , \Delta _i= \| x_i -x_{i-1} \| =(x_i-x_{i-1})^{1/N},
+\mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i}, 2 \leq i \leq k \right\} , \Delta _i= (x_i-x_{i-1})^{1/N},
 \end{equation}
-of the unknown H{\"o}lder constant $K$ from (\ref{eq4}) for the reduced function $\phi(x)$ where $r, r>1$ is a parameter of the algorithm; if $\mu=0$ accept $\mu=1$.
-\paragraph{Rule 3} Determine the best current value of the objective function
+of the unknown H{\"o}lder constant $K$ from (\ref{eq4}) for the reduced function $\phi(x)$, where $r>1$ is a parameter of the algorithm; if $\mu=0$ accept $\mu=1$.
+\paragraph{Rule 3.} Determine the best current value of the objective function
 \begin{equation}\label{eq8} 
@@
 
-\paragraph{Rule 4} For each interval $(x_{i-1},x_i),1 \leq i \leq k+1$ compute the value $R(i)$ called a \textit{characteristic} of the interval according to the formula
+\paragraph{Rule 4.} For each interval $(x_{i-1},x_i),1 \leq i \leq k+1$, compute the value $R(i)$ called a \textit{characteristic} of the interval according to the formula
 \begin{equation}\label{eq9} 
@@
   \begin{cases}
-    \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1}\\
-    2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1}\\
-    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1}
+    \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1},\\
+    2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1},\\
+    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
   \end{cases}
@@
 
-\paragraph{Rule 5} Select the interval $t$ with the largest value of characteristic $R(i):R(t)= \max_{1 \leq i \leq k+1} R(i)$.
+\paragraph{Rule 5.} Select the interval $t$ with the largest value of characteristic $R(i):R(t)= \max_{1 \leq i \leq k+1} R(i)$.
 
-\paragraph{Rule 6} Perform the next trial in the middle point of the interval $(x_{t-1},x_t)$ if the indices of its end points are not the same i. e.
+\paragraph{Rule 6.} Perform the next trial in the middle point of the interval $(x_{t-1},x_t)$, if the indices of its end points are not the same, i.e.
 \begin{equation}\label{eq10} 
-x^{k+1}=\frac {x_t+x_{t-1}}{2}, v(x_{t-1})\neq v(x_t),
+x^{k+1}=\frac {x_t+x_{t-1}}{2}, v(x_{t-1})\neq v(x_t).
 \end{equation}
@@
 \begin{equation}\label{eq11} 
-x^{k+1}= \frac {x_t+x_{t-1}}{2} -  \text{sign} {(z_t-z_{t-1})} \frac {{|z_t-z_{t-1}|}^N}{2r \mu ^N}, v(x_{t-1})=v(x_t).
+x^{k+1}= \frac {x_t+x_{t-1}}{2} -  \text{sign} {(z_t-z_{t-1})} \frac{1}{2r} \left[\frac {{|z_t-z_{t-1}|}}{\mu} \right]^N, v(x_{t-1})=v(x_t).
 \end{equation}
 
-\paragraph{Stop criterion} of the algorithm, according to which the execution of the algorithm is terminated for current global optimization problem (\ref{eq2}) is the achievement of a predefined accuracy of the problem solution i.
-e.
-if the length of the interval $(x_{t-1},x_t)$ containing current point $x^{k+1}$ is less or equal to the predefined accuracy $\Delta _t \leq \varepsilon$ where $t$ is from Rule 5 and $\varepsilon >0$ is the predefined accuracy.
+\paragraph{Stop criterion.} The search is terminated, if the length of the interval $(x_{t-1},x_t)$ containing current point $x^{k+1}$ is less than or equal to the predefined accuracy $\varepsilon$, i.e.
+$\Delta _t \leq \varepsilon$, where $t$ is from Rule 5 and $\varepsilon>0$ is the algorithm parameter.
 
 \subsection{ Approaches to the Reconstruction of the Objective Function Values }
-The problem of not everywhere defined objective function described above is a partial analog of the long-standing missing data problem in Data Science.
+
+The optimization problem with not everywhere defined objective function described above is a partial analog of the long-standing missing data problem in Data Science.
 The origin of this problem may be, for example, the technical problems or assembling the data set from several sources with different sets of parameters.
@@
 For example, LightGBM has a regime of ignoring the missing values while XGBoost restores the data by means of the decrease in the loss function when learning.
-Nevertheless, the majority of models cannot process the missing values and require the valid data without NaN or «missing» values.
+Nevertheless, the majority of models cannot process the missing values and require the valid data without NaN or ``missing'' values.
 It makes this problem to be serious enough.
@@
 \begin{itemize}
-  \item Deleting of all cases, which have missing values of some variables (the last resort, the algorithm loses the access to the useful information contained in the on-missing observations’ values);
+  \item Deleting of all cases, which have missing values of some variables (at the worst, the algorithm loses the access to the useful information contained in the on-missing observations’ values);
   \item Replacing any missing value with the average of the available values of the variable (doesn’t work with the qualitative variables, doesn’t account for the correlation of the data);
   \item Replacing by the most frequent value or by a constant (differs from the use of the median only by the ability to work with the qualitative parameters);
-  \item Replacing the data using the method of $k$ nearest neighbors or kNN (unlike the former, accounts for the correlation between the parameters, sensitive to the fadeouts, computation costly);
-  \item Multiple imputation by chained equations (MICE) – imputation of each value is performed not one but many times (allows understanding how reliable or unreliable is the value suggested);
-  \item Imputation of the data using a deep learning (restores the missing values by means of training a neural network on the points, which all parameters are available for, computation costly for big data).
+  \item Replacing the data using the method of $k$ nearest neighbors (unlike the former, accounts for the correlation between the parameters, sensitive to the fadeouts, computation costly);
+  \item Multiple imputation by chained equations (MICE) -- imputation of each value is performed not one but many times (allows understanding how reliable or unreliable is the value suggested);
+  \item Imputation of the data using a deep learning (restores the missing values by means of training a neural network on the points, for which all parameters are available, computation costly for big data).
 \end{itemize}
 
-The absence of an universal algorithm and the necessity to find the most suitable methods for each particular problem are the common drawbacks of the approaches listed above.
+The absence of a universal algorithm and the necessity to find the most suitable methods for each particular problem are the common drawbacks of the approaches listed above.
 
-For the imputation of the objective function values, one can consider, in general, a similar set of approaches, however, taking into account the Lipschitz condition and using the information-statistical theory.
-It allows selecting a universal algorithm fro this problem class.
-Next, the points, which the non-computability was found in are interpreted as the ones having the zero index, and (\ref{eq6}) is replaced by an extended formula
+For the imputation of the objective function values, one can consider, in general, a similar set of approaches, however, taking into account the Lipschitz condition and using the information-statistical approach.
+It allows to propose a good algorithm for this problem class.
+The points, at which the non-computability was found, are interpreted as the ones having the zero index, and (\ref{eq6}) is replaced by an extended formula
 \begin{equation}\label{eq12} 
@@
   \begin{cases}
-    -1 & {\quad \text{if } x_i \text{ is boundary point}}\\
-    1  & {\quad \text{if } x_i \text{ is internal point}}\\
-    0  & {\quad \text{if } x_i \text{ is non-computability}}
+    -1, & {\quad \text{if } x_i \text{ is boundary point}},\\
+    0, & {\quad \text{if } x_i \text{ is non-computable point}},\\
+    1, & {\quad \text{if } x_i \text{ is internal point}}.
   \end{cases}
@@
 
-Let us describe various possible approaches to the implementation of the modification.
+Let us describe various possible approaches to the imputation of the objective function values.
 
-\subsubsection{The first approach.}To ignore the absence of the objective function value in the point.
-The characteristic of an interval with even a single non-computable point is computed from the length of the interval
+\subsubsection{The first approach.} To ignore the absence of the objective function value at the point.
+The characteristic of an interval with even a single non-computable point is equal to the length of the interval, i.e.
 \begin{equation}\label{eq13} 
@@
 
-\subsubsection{The second approach.} To ignore the absence of the objective function values in the point.
-The characteristic of the interval with two non-computable point or with one non-computable point and one boundary point is computed from the interval multiplied by a reducing constant
+\subsubsection{The second approach.} To ignore the absence of the objective function values at the point.
+The characteristic of the interval with two non-computable point or with one non-computable point and one boundary point is equal to the interval length multiplied by a coefficient, i.e.
 \begin{equation}\label{eq14} 
-R(i)={(1- \frac {1}{r})}^N \Delta _i,
+R(i)={(1-{1}/{r})}^N \Delta _i,
 \end{equation}

### 2023-08-28 2023/MMST/samplepaper.tex
minor corrections

@@
 \includegraphics[width=\textwidth]{fig1.png}
-\caption{Examples of the distributions of the search trials in the non-computability regions in the first approach to the imputation of the values} \label{fig1}
+\caption{Examples of the search trials distribution in the non-computability regions for the first approach to the values imputation} \label{fig1}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{fig2.png}
-\caption{Examples of the distributions of the search trials in the non-computability regions in the second approach to the imputation of the values} \label{fig2}
+\caption{Examples of the search trials distribution in the non-computability regions for the second approach to the values imputation} \label{fig2}
 \end{figure}
 
-This approach resolves partly the problems of the first approach outlined above.
+This approach partly resolves the problems of the first approach outlined above.
 However, it doesn’t solve the problem of excess trials and requires a considerable increasing of the parameter $r$ for more complex test problem.
-The examples for the one-dimensional problems using the first two approaches are presented in ~\ref{fig1} and ~\ref{fig2} where the trials falling into the region with the undefined values are marked by the black points and the computable points of the search domain – by the green ones.
+The examples for the one-dimensional problems using the first two approaches are presented in Fig.~\ref{fig1} and Fig.~\ref{fig2} where the trials falling into the region with the undefined values are marked by the black points and the computable points of the search domain are marked by the green ones.
 
-\subsubsection{The third approach.} To restore the objective function values in the points of the intervals with two non-computable points as the averages of the neighbor points and to apply corresponding computation rule from formula (\ref{eq9}) assuming the points to be computable ones.
-If the neighbor points are non-computable as well, to use formula (\ref{eq14}).
-In the case of an interval with one non-computable point, to apply formulae (\ref{eq15}) and (\ref{eq16}).
+\subsubsection{The third approach.} To restore the objective function values at the points of the intervals with two non-computable points as the averages of the neighbor points and to apply corresponding computation rule from formula (\ref{eq9}) assuming the points to be computable ones.
+If the neighbor points are non-computable as well, use formula (\ref{eq14}).
+In the case of an interval with one non-computable point, apply formulae (\ref{eq15}) and (\ref{eq16}).
 \begin{figure}
@@
 
-Essential drawbacks of this approach are not utilizing the information on the coordinates aw well as seldom executing the computation rule added because of searching the substituting values among the nearest neighbors only.
+Essential drawbacks of this approach are not utilizing the information on the coordinates aw well as seldom executing the computation rule added because of searching the substituting values only among the nearest neighbors.
 
-\subsubsection{The fourth approach.} To restore the objective function values in the points for the intervals with two non-computable points as the averages of the nearest computable points taking into account the location coordinates i.
-e.
+\subsubsection{The fourth approach.} To restore the objective function values in the points for the intervals with two non-computable points as the averages of the nearest computable points taking into account the coordinates, i.e.
 \begin{equation}\label{eq17} 
@@
 where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval.
-The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located near the boundary point only.
+The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located only near the boundary point.
 \begin{figure}
@@
 \end{equation}
-This rule can be interpreted as an average value of the minorant and the majorant for a function satisfying the H{\"o}lder condition.
-Let us give the description of the modification of the global search algorithm in the case of not everywhere computable function based on this approach.
+The value $\tilde{z}(x)$ can be interpreted as an average value of the minorant and the majorant for a function satisfying the H{\"o}lder condition.
+Let us give the description of the modification of the global search algorithm in the case of partially defined function based on this approach.
 
-\subsubsection{Description of the modification} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions:
+\subsubsection{Description of the modification.} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions.
 \begin{enumerate} 
   \item The index $v_i=v(x_i)$ is determined by the formula (\ref{eq12}).
-  \item When computing current lower estimate of the H{\"o}lder constant $K$ in formula (\ref{eq7}), use the intervals with two computable points only.
-  \item The best current objective function value is determined for different values of the index
+  \item When computing current lower estimate of the H{\"o}lder constant $K$ in formula (\ref{eq7}), use only the intervals with two computable boundary points.
+  \item The value $z_v^*$ depends on the index value
 \begin{equation}\label{eq20} 
@@
 \end{equation}
-where $\varepsilon _r$ is a parameter of the method.
+where $\varepsilon _r > 0$ is a parameter of the method.
 

### 2023-08-29 2023/MMST/samplepaper.tex
fix terminology

@@
 This paper presents the results in a new research direction related to the use of imputation\footnote{The term ``imputation'' in the paper is used following the established terminology in the field of machine learning related to the restoring of the missing data.} (restoring) of the objective function values in the course of the global search algorithm operation when solving the problems with partially defined objective functions.
-The approach proposed is based on an adoption of the computation rules of the global search algorithm.
+The approach proposed is based on an adaption of the computation rules of the global search algorithm.
 The efficiency of the approach proposed has been demonstrated numerically.
@@
 
-  \item For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$, the computing of the characteristic $R(i)$ is performed according to the following computing formulae.
+  \item For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$, the computing of the characteristic $R(i)$ is performed according to the following formulae.
 
-        \paragraph{Case 1} For the intervals with \textbf{two computable points} or with \textbf{one computable point and one boundary point}, apply the classic rules of global search algorithm
+        \paragraph{Case 1.} For the intervals with \textbf{two computable points} or with \textbf{one computable point and one boundary point}, apply the original rules of the global search algorithm
 \begin{equation}\label{eq21} 
@@
   \begin{cases}
-     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z_1^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1}\\
-    2 \Delta _i-4 \frac {(z_i-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1}\\
-    2 \Delta _i-4 \frac {(z_{i-1}-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1}
+     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z_1^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1,}\\
+    2 \Delta _i-4 \frac {(z_i-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1,}\\
+    2 \Delta _i-4 \frac {(z_{i-1}-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
   \end{cases}
 \end{equation}
-        \paragraph{Case 2}   For the intervals with \textbf{one boundary point and one non-computable point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
+        \paragraph{Case 2.}   For the intervals with \textbf{one boundary point and one non-computable point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
 \begin{equation}\label{eq22} 
@@
   \begin{cases}
-    2 \Delta _i-4 \frac {(z_{i+1}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=0}\\
-    2 \Delta _i-4 \frac {(z_{i-2}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=-1}
+    2 \Delta _i-4 \frac {(z_{i+1}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=0,}\\
+    2 \Delta _i-4 \frac {(z_{i-2}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=-1.}
   \end{cases}
 \end{equation}
-        \paragraph{Case 3} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor computable point and apply the formula
+        \paragraph{Case 3.} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor computable point and apply the formula
 \begin{equation}\label{eq23} 
@@
   \begin{cases}
-     \Delta _i+\frac {{(z_i-z_{i-2})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-2}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=1}\\
-     \Delta _i+\frac {{(z_{i+1}-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_{i+1}+z_{i-1}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0}
+     \Delta _i+\frac {{(z_i-z_{i-2})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-2}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=1,}\\
+     \Delta _i+\frac {{(z_{i+1}-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_{i+1}+z_{i-1}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0.}
   \end{cases}
 \end{equation}
-        \paragraph{Case 4} For the intervals with \textbf{two non-computable points}, perform the imputation of the values for the non-computable points based on the rule (\ref{eq19}) and apply the formula
+        \paragraph{Case 4.} For the intervals with \textbf{two non-computable points}, perform the imputation of the values for the non-computable points based on the rule (\ref{eq19}) and apply the formula
 \begin{equation}\label{eq24} 
-R(i)=\Delta _i+\frac {{(\tilde{z}_i-\tilde{z}_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {\tilde{z}_i+\tilde{z}_{i-1}-2z_1^*}{r \mu}, v(x_{i-1})=v(x_i)=0
+R(i)=\Delta _i+\frac {{(\tilde{z}_i-\tilde{z}_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {\tilde{z}_i+\tilde{z}_{i-1}-2z_1^*}{r \mu}, \; v(x_{i-1})=v(x_i)=0.
 \end{equation}
-        \paragraph{Case 5} In the case when any imputation is impossible, use the formula
+        \paragraph{Case 5.} In the case when any imputation is impossible, use the formula
 \begin{equation}\label{eq25} 
@@
 \end{equation}
-  \item If several intervals have equal characteristics, prefer the computable interval with the lowest number.
+  \item If several intervals have equal characteristics, prefer the interval with computable boundary points that has the lowest number.
   \item If the best interval appears to be the one with two non-computable points, perform the next trial in the middle point of the interval.
@@
 
-\subsection{Results of Numerical Experiments}
-The numerical experiments were carried out using a computer with Intel® Core™ i7-10750H CPU @ 2.60GHz 2.59 GHz using the developed software package implementing the modification described in Section 3.
+\section{Results of Numerical Experiments}
+
+The numerical experiments were carried out using a computer with Intel® Core™ i7-10750H CPU @ 2.60GHz using the developed software package implementing the modification described in Section 3.
 To generate a series of problems, GKLS generator described in \cite{Gaviano2003} was used.
-It allows generating the multiextremal oprimization problems with the properties known in advance.
+It allows generating the multiextremal optimization problems with the properties known in advance.
 These problems were appended with the non-computable regions of various kinds.
 
-Each series of experiments consisted of 100 GKLS problems with everywhere computable objective function (the classic GKLS problems), 100 problems with the non-computabilities at the boundaries of the search domains, and 100 problems with non-computability regions defined randomly.
+Each series of experiments consisted of 100 GKLS problems with everywhere computable objective function (the original GKLS problems), 100 problems with the non-computabilities at the boundaries of the search domains, and 100 problems with non-computability regions defined randomly.
 
@@
 The problems of the dimensionality $N=2$ were used in the experiment.
-The reliability parameter was set to $r=5.5$, the precision $\varepsilon=0.01$.
-The results of experiments are presented in Tab.~\ref{tab1}.
+The reliability parameter was set to $r=5.5$, the accuracy was $\varepsilon=0.01$.
+The results of experiments are presented in Table~\ref{tab1}.
 
-\begin{figure}
-\includegraphics[width=\textwidth]{fig6.png}
-\caption{Examples of problems \#\# 19, 25, and 35 from the first series of experiments} \label{fig6}
-\end{figure}
-
-Fig.~\ref{fig6} presents the examples of the modification operation with several problems of the series.
-The trials inside the computable regions are denoted by the blue points in the figures, the global minima of the problems – by the red points, and the points of trials performed inside the non-computability regions – by black.
-Also, the non-computability regions are filled with grey for clarity.
-
-Since the modification works with the classic GKLS problem series according to the base global search algorithm, one can compare it with the classic GSA.
-Thus, the general behavior of GSA was preserved for the modification, and the same regions of trials’ densification near the local minima were observed that can be explained by the realization of the classical computation rules (\ref{eq21}) in most cases.
-One can see in Tab.~\ref{tab1} that the number of trials executed inside the non-computable regions is small.
-Total number of trials for the classic problems and for the problems with non-computable regions are comparable, the method doesn’t perform excess iterations in the non-promising regions due to the computing formulae utilizing the information on the nearest computable points.
-
-\begin{table}
-\caption{Comparison of the efficiencies of the algorithm modification with the classic version of the global search algorithm on GKLS problems}\label{tab1}
+\begin{table}[h]
+\caption{Comparison of the original and modified GSA}\label{tab1}
 \begin{tabular}{|l|c|c|c|}
@@
 \hline
-Average number of iterations & 781,66	& 854,91 & 839,03 \\
-Average number of non-computability & - & 185,79 & 42,31 \\
+Average number of trials & 781,66	& 854,91 & 839,03 \\
+Average number of trials in non-computability regions & - & 185,79 & 42,31 \\
 Number of solved problems & 100 / 100 & 100 / 100 & 100 / 100\\
@@
 
+\begin{figure}[h]
+\includegraphics[width=\textwidth]{fig6.png}
+\caption{Examples of problems \#\# 19, 25, and 35 from the first series of experiments} \label{fig6}
+\end{figure}
+
+Fig.~\ref{fig6} presents the examples of solving several problems with the modified GSA.
+The trials inside the computable regions are marked by the blue points, the global minima of the problems are marked by the red points, and the points of trials performed inside the non-computability regions are marked by the black points.
+Also, the non-computability regions are filled with grey for clarity.
+
+Since the modified algorithm works with the original GKLS problem series according to the rules of base algorithm, one can compare it with the GSA.
+Thus, the general behavior of GSA was preserved for the modification, and the same regions of trials' accumulation near the local minima were observed that can be explained by the realization of the  computation rules (\ref{eq21}) in most cases.
+One can see in Table~\ref{tab1} that the number of trials executed inside the non-computable regions is small.
+Total number of trials for the original problems and for the problems with partially defined objective function are comparable, the method doesn't perform excess trials in the non-promising regions due to the computing formulae utilizing the information on the nearest computable points.
+
 The use of rule (\ref{eq25}) forms a grid of trials inside the non-computability regions that is visible clearly for Problem \#35.
@@
 100\% of the problems were solved.
-However, for the series with the boundary non-computability region, the value of parameter $r=6.0$ was required to solve one problem, for the series with the random regions – soving of two problems with given value of parameter $r$.
+However, for the series with the boundary non-computability region the value of parameter $r=6.0$ was required to solve one problem, and for the series with the random non-computability regions the same value of $r$ was required to solve two problems.
 
-In the second series of experiments, the problem series of various dimensionalities were solved.
-In the experiments with the problems of the dimensionality $N=4$, the reliability parameter was set to $r=4.5$, the precision $\varepsilon=0.02$.
+In the second series of experiments, the problem series of various dimensions were solved.
+In the experiments with the problems of the dimensions $N=3,4,5$, the reliability parameter was set to $r=4.5$, the precision $\varepsilon=0.02$.
 The global minimum was considered to be found if the algorithm generates a trial point inside the $\varepsilon$-nearness of the global minimum.
-The results of the experiments are presented in Tab.~\ref{tab2}.
+The results of the experiments are presented in Table~\ref{tab2}.
 
 \begin{table}
-\caption{Problem series of various dimensionalities}\label{tab2}
+\caption{Problem series of various dimensions}\label{tab2}
 \begin{tabular}{|l|l|c|c|c|}
 \hline
- N &  Task type & Average num. of iters & Average num. of non-comput. & Solved problems \\
+ N &  Task type & Average num. of trials & Average num. of non-comput. & Solved problems \\
 \hline
-$3$ & classical & 12251,11 & - & 100 / 100 \\
-$ $ & round & 12201,78 & 4763,57 & 100 / 100 \\
-$ $ & random & 12210,71 & 63,39 & 100 / 100 \\
+$3$ & classical & 12251.1 & - & 100 / 100 \\
+$ $ & round & 12201.8 & 4763.6 & 100 / 100 \\
+$ $ & random & 12210.7 & 63.4 & 100 / 100 \\
 \hline
-$4$ & classical & 25144,46 & - & 100 / 100 \\
-$ $ & round & 29840,32 & 15004,37  & 100 / 100 \\
-$ $ & random & 25287,49 & 26,43 & 100 / 100 \\
+$4$ & classical & 25144.5 & - & 100 / 100 \\
+$ $ & round & 29840.3 & 15004.4  & 100 / 100 \\
+$ $ & random & 25287.5 & 26.4 & 100 / 100 \\
 \hline
-$5$ & classical & 56725,95 & - & 100 / 100 \\
-$ $ & round & 47840,09 & 24109,19 & 100 / 100\\
-$ $ & random & 57874,76 & 82,57 & 100 / 100\\
+$5$ & classical & 56725.9 & - & 100 / 100 \\
+$ $ & round & 47840.1 & 24109.2 & 100 / 100\\
+$ $ & random & 57874.8 & 82.6 & 100 / 100\\
 \hline
@@
 
-According to the results of the experiments, the method resolves successfully the occasional falls into the non-computability regions scattered inside the search domain randomly whereas in the case of large boundary regions it tries the execute a complete investigation without executing too many excess trials.
+The results of the experiments show that the method resolves successfully the hits into the non-computability regions randomly scattered inside the search domain whereas in the case of large boundary regions it tries the execute a complete exploration without executing too many excess trials.
 
-\subsection{Conclusions}
-In the present work, the problem of finding the global minimum of a computation-costly "black box"\ function was considered.
-As compared to the traditional formulation used in the global optimization, the problem considered has an important difference: the objective function may be undefined in some regions of the search domain (in particular, in one point or in several ones).
-The appearing of such a problem statement is caused by a frequent problem of numerical instability of the models investigated in the applied problems.
+\section{Conclusions}
 
-The main problem is the impossibility to identify the non-computability regions in advance in most cases.
-In fact, the region of feasible combinations of parameters is undefined a priori.
-This phenomenon can be interpreted as the presence of additional hidden constraints in the problem.
+In the present work, the problem of finding the global minimum of a computation-costly ``black box'' function was considered.
+As compared to the traditional formulation used in the global optimization, the considered problem has an important difference: the objective function may be undefined in some regions of the search domain.
+The appearing of such a formulation is caused by a frequent problem of numerical instability of the models investigated in the applied tasks.
+
+The main problem is the impossibility to identify the non-computability regions in advance in the most cases.
+In fact, the region of feasible combinations of parameters is unknown a priori.
+This phenomenon can be interpreted as partial computability of the objective function.
 
@@
 As a solution, the experts in the field have proposed various approaches to the imputation (restoring) of the missing values.
-These ideas have found their reflection in the development of the new method.
+%These ideas have found their reflection in the development of the new method.
+The applicability of already existing approaches appended with the different variants to account for the Lipschitz condition was investigated.
+As a result, efficient algorithm for solving the problems of this class has been proposed.
+This algorithm is a modification of the deterministic method for solving the Lipschitz global optimization problems, namely information-statistical global search algorithm.
+%Within the framework of this modification, the scheme of the algorithm were adopted with introduction of the data imputation rules.
+As part of the modification, the computational scheme of the algorithm was adapted with the introduction of the data imputation rules.
 
-The applicability of already existing approaches appended with various variants to account for the Lipschitz condition was investigated.
-As a result, a universal efficient algorithm for solving the problems of this class has been developed.
-The algorithm constructed is a development of one of efficient deterministic method of solving the Lipschitz global optimization problems, namely information-statistical global search algorithm.
-Within the framework of this modification, the computation rules of the algorithm were adopted with introduction of the data imputation algorithm.
-
-The numerical experiments were carried out using series of multiextremal GKLS problems of the dimensionalities $N = 2,3,4,5$.
-The results have demonstrated the reliability and efficiency of the approach proposed.
-The modified algorithm handles successfully the problems, which the non-computability regions are scattered randomly within the search domain.
-In the case of large boundary non-computability regions, the method solves the problems efficiently as well performing a complete investigation of the non-computability regions also.
+The numerical experiments were carried out using series of multiextremal GKLS problems of the dimensions $N = 2,3,4,5$.
+The results have demonstrated the reliability and efficiency of the proposed approach.
+The modified algorithm handles successfully the problems, in which the non-computability regions are randomly scattered within the search domain.
+In the case of large boundary non-computability regions, the method solves the problems efficiently as well performing an exploration of the non-computability regions.
 

### 2023-08-29 2023/MMST/samplepaper.tex
add some fix

@@
 \includegraphics[width=\textwidth]{fig3.png}
-\caption{Imputation of the non-computable values by the averages of the neighbor points. Approach \#3} \label{fig3}
+\caption{Imputation of the non-computable values by the averages of the neighbor points. Approach No. 3} \label{fig3}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{fig4.png}
-\caption{Imputation of non-computable values by averages of the neighboring points. Approach \#4} \label{fig4}
+\caption{Imputation of non-computable values by averages of the neighboring points. Approach No. 4} \label{fig4}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{fig5.png}
-\caption{Imputation of non-computable values by averages of the minorant and the majorant. Approach \#5} \label{fig5}
+\caption{Imputation of non-computable values by averages of the minorant and the majorant. Approach No. 5} \label{fig5}
 \end{figure}
@@
 
-\subsubsection{Description of the modification.} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions.
+\subsubsection{Description of the modified algorithm.} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions.
 \begin{enumerate} 
@@
 
-        \paragraph{Case 1.} For the intervals with \textbf{two computable points} or with \textbf{one computable point and one boundary point}, apply the original rules of the global search algorithm
+        \paragraph{Case 1.} For the intervals with \textbf{two computable points} or with \textbf{one computable and one boundary point}, apply the original rules of the global search algorithm
 \begin{equation}\label{eq21} 
@@
 \end{equation}
-        \paragraph{Case 2.}   For the intervals with \textbf{one boundary point and one non-computable point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
+        \paragraph{Case 2.}   For the intervals with \textbf{one non-computable and one boundary point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
 \begin{equation}\label{eq22} 
@@
 
-The numerical experiments were carried out using a computer with Intel® Core™ i7-10750H CPU @ 2.60GHz using the developed software package implementing the modification described in Section 3.
+The numerical experiments were carried out using a computer with Intel\textregistered\ Core\texttrademark\ i7-10750H CPU @ 2.60GHz using the developed software package implementing the modification described in Section 3.
 To generate a series of problems, GKLS generator described in \cite{Gaviano2003} was used.
@@
 
+\begin{figure}[h]
+\includegraphics[width=\textwidth]{fig6.png}
+\caption{Examples of problems No.No. 19, 25, and 35 from the first series of experiments} \label{fig6}
+\end{figure}
+
 \begin{table}[h]
 \caption{Comparison of the original and modified GSA}\label{tab1}
+\centering
 \begin{tabular}{|l|c|c|c|}
 \hline
- &  classical  &  round district  &  random district  \\
+ &  { classical }  & { round district }  & { random district }  \\
 \hline
-Average number of trials & 781,66	& 854,91 & 839,03 \\
-Average number of trials in non-computability regions & - & 185,79 & 42,31 \\
-Number of solved problems & 100 / 100 & 100 / 100 & 100 / 100\\
+Average number of trials & 781.7	& 854.9 & 839.0 \\
+\begin{tabular}{@{}l@{}} {Average number of trials } \\  { in non-computability regions } \end{tabular}  & - & 185.8 & 42.3 \\
+{Number of solved problems } & 100 / 100 & 100 / 100  &  100 / 100 \\
 \hline
@@
 \end{table}
-
-\begin{figure}[h]
-\includegraphics[width=\textwidth]{fig6.png}
-\caption{Examples of problems \#\# 19, 25, and 35 from the first series of experiments} \label{fig6}
-\end{figure}
 
@@
 
-The use of rule (\ref{eq25}) forms a grid of trials inside the non-computability regions that is visible clearly for Problem \#35.
+The use of rule (\ref{eq25}) forms a grid of trials inside the non-computability regions that is visible clearly for Problem No.
+35.
 At that, the trial points are denser at the boundary of the non-computable region close to potential minimum when attempting to approach this one from different sides due to the computational rules (\ref{eq22})-(\ref{eq24}) that was observed in all problems presented as the examples.
@@
 
-In the second series of experiments, the problem series of various dimensions were solved.
-In the experiments with the problems of the dimensions $N=3,4,5$, the reliability parameter was set to $r=4.5$, the precision $\varepsilon=0.02$.
-The global minimum was considered to be found if the algorithm generates a trial point inside the $\varepsilon$-nearness of the global minimum.
-The results of the experiments are presented in Table~\ref{tab2}.
-
 \begin{table}
 \caption{Problem series of various dimensions}\label{tab2}
+\centering
 \begin{tabular}{|l|l|c|c|c|}
 \hline
- N &  Task type & Average num. of trials & Average num. of non-comput. & Solved problems \\
+ N &  Task type & Average number of trials & \begin{tabular}{@{}c@{}} {Average number of trials} \\  {in non-computability regions} \end{tabular} & Solved problems \\
 \hline
@@
 \end{table}
+
+In the second series of experiments, the problem series of various dimensions were solved.
+In the experiments with the problems of the dimensions $N=3,4,5$, the reliability parameter was set to $r=4.5$, the precision $\varepsilon=0.02$.
+The global minimum was considered to be found if the algorithm generates a trial point inside the $\varepsilon$-nearness of the global minimum.
+The results of the experiments are presented in Table~\ref{tab2}.
 

### 2023-08-29 2023/MMST/samplepaper.tex
add more references

--- 2023-08-29 2023/MMST/samplepaper.tex
@@
 Often such information is represented in the form of suggestion that the objective function $\phi$ satisfies the Lipschitz condition with a priori unknown constant $L$.
-
 This suggestion can be interpreted (in the relation to the applied problems) as the reflection of a limited power generating the changes in the simulated system.
@@
 
-A number of efficient deterministic algorithms for solving the Lipschitz global optimization problems have been developed ~\cite{Jones2009,Birect2020,Sergeyev2017}.
-The comparisons conducted have shown the deterministic algorithms to overcome (in various criteria) the widespread nature-inspired algorithms ~\cite{Liberti2005,Sergeyev2018,Sovrasov2019} .
+A number of efficient deterministic algorithms for solving the Lipschitz global optimization problems have been developed ~\cite{Grishagin2018,Jones2021,Birect2020,PaulaviciusZilinskas2014,Sergeyev2017}.
+The comparisons conducted have shown the deterministic algorithms to overcome (in various criteria) the widespread nature-inspired algorithms ~\cite{Liberti2005,Sergeyev2018}.
 
-The present work continues the development of one of the efficient deterministic algorithms for solving the Lipschitz global optimization problems, namely the of the information-statistical global search algorithm ~\cite{indexMethod,strongin1978,Strongin2000}.
-This algorithm suggests the solving of a multidimensional problem by means of its reduction to an equivalent one-dimensional optimization problem using the Peano-Hilbert space-filling curves \cite{Sergeyev2013}.
+The present work continues the development of one of the efficient deterministic algorithms for solving the Lipschitz global optimization problems, namely the of the information-statistical global search algorithm ~\cite{Strongin2000,Sergeyev2013}.
+This algorithm suggests the solving of a multidimensional problem by means of its reduction to an equivalent one-dimensional optimization problem using the Peano-Hilbert space-filling curves.
 
 Note that the use of complex mathematical models is natural in the optimization of complex real world objects that, as a consequence, essentially increases the computation costs for searching the optimum.
-In last decades, the experts in the fields of optimization and parallel computations have proposed many methods for reducing the computation costs an speed-up of the algorithms related to solving the appearing optimization problems \cite{Kvasov2013}  as well as to the numerical analysis of the original models \cite{Dongarra2022,Duwe2020}.
+In last decades, the experts in the fields of optimization and parallel computations have proposed many methods for reducing the computation costs an speed-up of the algorithms related to solving the appearing optimization problems \cite{Kvasov2013,Sergeyev2020} as well as to the numerical analysis of the original models \cite{Dongarra2022,Duwe2020}.
 
@@
 In such regions, one cannot perform the numerical modeling and compute the objective function values correctly.
-This phenomenon can be interpreted either as the presence of some hidden (unknown a priori) constraints or as a partial computability of the objective function in the search domain.
-In such a statement, the optimization problem becomes more complex since the region of feasible parameters’ combinations is not defined in advance.
+This phenomenon can be interpreted either as the presence of some hidden (unknown a priori) constraints \cite{Stripinis2021} or as a partial computability of the objective function in the search domain \cite{Candelieri2019,Sergeyev2003,Strongin2020}.
+In such a statement, the optimization problem becomes more complex since the region of feasible parameters' combinations is not defined in advance.
 
@@
 
-\subsection{Description of the Global Search Algorithm}
+\subsection{Global Search Algorithm}
 
-Within the framework of the proposed approach, the information-statistical theory of global search is used for solving the problems (\ref{eq2}).
+%Within the framework of the proposed approach, the information-statistical theory of global search is used for solving the problems (\ref{eq2}).
 The main idea of the algorithm is the use of the accumulated information to determine the next interval, within which it is most probable to find the absolute minimum.
@@
 
-Let us give more detailed description of the algorithm.
+Let us give more detailed description of the \textbf{global search algorithm} (GSA) for solving the problem (\ref{eq2}).
 
-\subsubsection{Global search algorithm (GSA)} At every iteration of the global search a \textit{trial} is executed.
-We will refer to computing the objective function value  $\phi (x)$ from (\ref{eq2}) as a trial.
+At every iteration of the global search a \textit{trial} is executed.
+We will refer to computing the objective function value  $\phi (y(x))$ from (\ref{eq2}) as a trial.
 The first trial will be executed in the middle point $x^1 \in (0,1)$.
@@
 
-A large number of excess trials after getting into the non-computability region due to a large value of the characteristic obtained according to (\ref{eq13}) is an essential draw-back of this approach.
+A large number of excess trials after getting into the non-computability region due to a large value of the characteristic obtained according to (\ref{eq13}) is an essential drawback of this approach.
 
@@
 \begin{equation}\label{eq17} 
-\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
+%\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
+\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \rho(x,x')}{\rho(x',x'') },
 \end{equation}
-where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval.
+where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval and $\rho(x_1,x_2) =  |x_1 - x_2|^{1/N}$.
 The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located only near the boundary point.
@@
     z', & {\quad x \leq \hat{x}',}\\
-    \frac {z'+z''}{2}- \frac {r \mu}{2} (\| x - x' \| + \| x - x'' \|), & {\quad \hat{x}' < x < \hat{x}'',}\\
-    z'',  & {\quad x \geq \hat{x}''.}
+    %\frac {z'+z''}{2}- \frac {r \mu}{2} (\| x - x' \| + \| x - x'' \|), & {\quad \hat{x}' < x < \hat{x}'',}\\
+    \frac {z'+z''}{2}- \frac {r \mu}{2} (\rho(x,x')  + \rho(x,x'')), & {\quad \hat{x}' < x < \hat{x}'',}\\
+    z'',  & {\quad x \geq \hat{x}'',}
   \end{cases}
 \end{equation}
+where $\rho(x_1,x_2) =  |x_1 - x_2|^{1/N}$.
 The value $\tilde{z}(x)$ can be interpreted as an average value of the minorant and the majorant for a function satisfying the H{\"o}lder condition.
@@
 \end{equation}
-        \paragraph{Case 2.}   For the intervals with \textbf{one boundary point and one non-computable point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
+        \paragraph{Case 2.}   For the intervals with \textbf{one non-computable point and one boundary point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
 \begin{equation}\label{eq22} 
@@
 \includegraphics[width=\textwidth]{fig6.png}
-\caption{Examples of problems \#\# 19, 25, and 35 from the first series of experiments} \label{fig6}
+\caption{Examples of problems No. 19, 25, and 35 from the first series of experiments} \label{fig6}
 \end{figure}

### 2023-08-29 2023/ITMO paper/elsarticle-template-num.tex

@@
 
-На качество работы методов машинного обучения существенное влияние оказывают их гиперпараметры, при этом оценка критерия качества является трудоемкой операцией.
-Поэтому актуальной является задача разработки ителлектуальных методов для выбора оптимальных значений гиперпараметров, которые требуют малого числа поисковых испытаний.
-В этой статье мы предлагаем новый подход к настройке гиперпараметров, основанный на идеях Lipschitz global optimization. 
-В рамках данного подхода решение задач с несколькими параметрами сводится к решению эквивалентных им одномерных задач.
-Соответствующая редукция основана на использовании space-filling curves (Peano curves).
-Указанные подходы реализованы в open-source фреймворке методов интеллектуальной оптимизации iOpt.
-Для демонстрации преимуществ iOpt проведено его сравнение с известными фреймворками Optuna и HyperOpt при настройке гиперпараметров различных методов машинного обучения на представительном наборе датасетов.
-Результаты показывают, что методы Lipschitz global optimization обеспечивают сравнимые по качеству результаты за существенно меньшее время по сравнению с известными алгоритмами настройки гиперпараметров.
+The quality of machine learning methods is substantially affected by their hyperparameters, while the evaluation of the quality criterion is a time-consuming operation.
+Therefore, it is important to develop intelligent methods for selecting optimal values of hyperparameters that require a small number of search trials.
+In this paper, we propose a new approach to hyperparameter tuning based on the ideas of Lipschitz global optimization.
+In the framework of this approach, the solution of problems with several parameters is reduced to solving equivalent one-dimensional problems.
+The reduction is based on the use of space-filling curves (Peano curves).
+These approaches are implemented in the iOpt open-source framework of intelligent optimization methods.
+To demonstrate the advantages of iOpt, we compare it with the well-known Optuna and HyperOpt frameworks when tuning hyperparameters of various machine learning methods on a representative set of datasets.
+The results show that Lipschitz global optimization methods provide comparable (in terms of quality) results in a significantly shorter time compared to known hyperparameter tuning algorithms.
 
@@
 \begin{highlights}
-\item Для настройки гиперпараметров разработан фреймворк методов липшицевой оптимизации
-\item Для решения многомерных задач используется редукция размерности на оснвое кривых Пеано
-\item Предложена схема сведения задач, часть параметров которых является категориальными, к одной задаче с непрерывными параметрами 
-\item Разработанные методы (в отличие от известных методов настройки гиперпараметров) являются детерминированными, что обеспечивает стабильный результат при разных запусках алгоритмов
+\item A framework of Lipschitz optimization methods is developed for the tuning of hyperparameters
+\item Dimensionality reduction based on Peano space-filling curves is used to solve multidimensional problems
+\item A scheme is proposed for reducing the problems, some of whose parameters are categorical, to a single problem with continuous parameters
+\item The developed methods (unlike the known methods of hyperparameter tuning) are deterministic, which ensures stable results at different runs of the algorithms
+
 \end{highlights}
@@
 
-В настоящее время сложные научно и/или промышленные задачи чаще всего решаются численными методами.
-В большинстве таких методов, например, в методах машинного обучения, имеются гиперпараметры, от выбора конкретных значений которых может зависеть эффективность (в заданной метрике) полученного решения при заданных тем или иным способом ограничениях на время расчетов \cite{Hutter2019,nikitin2021}.
-Тот факт, что во многих случаях разница в эффективности получаемого решения при разных значениях гиперпараметров может быть достаточно значительной, привел к возникновению the hyperparameter optimization (HPO) problem, а также исследованиям подходов к ее решению, реализуемых в фреймворках автоматического подбора гиперпараметров \cite{Tune, optuna, hyperopt,Sherpa}.
-С математической точки зрения HPO problem соответствует задаче глобальной оптимизации с фиксированными границах изменения гиперпараметров.
-В некоторых случаях часть гиперпараметров может быть категориальной, то есть принимать значения лишь из некоторого, обычно небольшого, числа вариантов, что приводит к необходимости использовать методы оптимизации, способные решать задачи с дискретными параметрами.
+Currently, complex scientific and/or industrial problems are most often solved by numerical methods.
+Most of such methods, for example, machine learning methods, have hyperparameters, and the choice of specific values of these hyperparameters can determine the efficiency (in a given metric) of the obtained solution under certain constraints on the computation time \cite{Hutter2019,nikitin2021}.
+The fact that in many cases the difference in the efficiency of the solution obtained with different values of hyperparameters can be quite significant, has resulted in the emergence of the hyperparameter optimization (HPO) problem, as well as the research of approaches to its solution that are implemented in the frameworks for automatic selection of hyperparameters \cite{Tune, optuna, hyperopt,Sherpa}.
+From the mathematical point of view, the HPO problem corresponds to a global optimization problem with fixed bounds on the variation of hyperparameters.
+In certain cases, some of the hyperparameters may be categorical, i.e., they can only assume values from a certain, usually small, number of variants, which necessitates the use of optimization methods capable of solving problems with discrete parameters.
 
-Для конечного пользователя, желающего провести настройку используемого метода, подобрав оптимальное сочетание гиперпараметров, выбор конкретного фреймворка определяется следующими факторами: 1) эффективность реализованных в фреймворке методов, позволяющих решить HPO problem того класса задач, которому принадлежит задача пользователя; 2) простота использования фреймворка с точки зрения описания/формирования задачи подбора оптимального сочетания гиперпараметров, организации процесса поиска оптимума, а также получения и анализа результатов.
-В настоящее время наиболее распространенным способом организации такого интерфейса является использование языка программирования Python, при этом сами фреймворки могут быть написаны, как на этом, так и на других языках.
+For an end user who wishes to fine-tune the method used by selecting an optimal combination of hyperparameters, the choice of a particular framework is determined by the following factors: 1) efficiency of the methods implemented in the framework that allow solving an HPO problem from the class of problems to which the user's problem belongs; 2) ease of using the  framework from the point of view of description/formulation of the problem of selecting the optimal combination of hyperparameters, organization of the optimum search process, as well as obtaining and analyzing the results.
+Today, the most common way to organize such an interface is to use the Python programming language, while the frameworks can be written both in this and other languages.
 
-В данной статье мы представляем фреймворк методов интеллектуальной оптимизации iOpt\footnote{\url{https://github.com/aimclub/iOpt/}}, ориентированный на решение задач выбора оптимальных значений гиперпараметров различных алгоритмов (включая методы машинного обучения и эвристической оптимизации).
-Статья построена следующим образом.
-В начале приведен обзор известных методов, используемых для настройки гиперпараметров.
-В section \ref{sec_lip} приведена математическая постановка задачи, описаны одходы к редукции многомерных задач к одномерным и решению задач с дискретными параметрами.
-В section \ref{sec_iOpt} описана алгоритмическая основа методов глобальной оптимизации, реализованных во фреймворке iOpt.
-В section \ref{sec_exp} даны результаты численных экспериментов, проведенных для сравнения эффективности фреймворка iOpt с другими известными фреймворками на модельных и прикладных задачах.
-В заключении приведены выводы и планы по развитию фреймворка.
+In this paper, we present iOpt \footnote{\url{https://github.com/aimclub/iOpt/}}, a framework of intelligent optimization methods focused on solving problems of selecting optimal values of hyperparameters of various algorithms (including machine learning and heuristic optimization methods).
+The paper is organized as follows.
+At the beginning, there is a review of known methods used for hyperparameter tuning.
+Section \ref{sec_lip} presents the mathematical formulation of the problem, describes approaches to the reduction of multidimensional problems to one-dimensional problems and to the solution of problems with discrete parameters.
+Section \ref{sec_iOpt} explains the algorithmic basis of global optimization methods implemented in the iOpt framework.
+In Section \ref{sec_exp}, we present the results of numerical experiments conducted to compare the efficiency of the iOpt framework with other known frameworks on model and applied problems.
+The conclusion summarizes the findings and plans   for the development of the framework.
 
@@
 
-Основная сложность при решении HPO problem состоит в том, что поиск оптимального сочетания гиперпараметров требует для каждого выбранного варианта их значений решения соответствующей исходной задачи, например, задачи машинного обучения, а значит, может быть довольно длительным по времени.
-Таким образом, любой подход к решению HPO problem будет ограничен в числе сочетаний гиперпараметров, которое метод глобальной оптимизации сможет проверить прежде, чем будет исчерпан доступный вычислительный ресурс.
-Конечно, в таких условиях сложно говорить о сходимости метода к глобальному оптимуму или о каких-то гарантиях его отыскания.
-Речь может идти только о возможности получения приемлемого (с точки зрения исследователя) решения.
-Дадим краткую характеристику ряда методов, которые используются для задач настройки гиперпараметров (в зависимости от свойств исходно задачи).
+The main difficulty in solving the HPO problem is that the search for the optimal combination of hyperparameters requires for each selected variant of their values to solve the corresponding initial problem, for example, the problem of machine learning, and thus can take quite a long time.
+Thus, any approach to solving the HPO problem will be limited by the number of combinations of hyperparameters that the global optimization method can test before the available computational resource is exhausted.
+Certainly, it is difficult in such circumstances to speak about convergence of the method to the global optimum or about any guarantees of finding it.
+We can only consider the possibility of obtaining an acceptable (from the researcher's point of view) solution.
+Let us briefly characterize a number of methods that are used to tune hyperparameters (depending on the properties of the initial problem).
 
-Самым простым способом поиска глобального оптимума является метод полного перебора (поиска по равномерной \cite{Bao2006} или случайной \cite{Bergstra2012} сетке).
-Выбрав число анализируемых вариантов по каждому из гиперпараметров, перебрав все получившиеся сочетания и взяв наилучшее из них в заданной метрике, мы получим оценку оптимума с определенной точностью \cite{Nevendra2022}.
-Метод полного перебора прост в реализации, обладает идеальным параллелизмом, но имеет существенный недостаток – экспоненциальный рост числа узлов сетки с ростом числа гиперпараметров.
-Таким образом, в качестве метода для решения HPO problem полный перебор может использоваться только в «простых» задачах, где малО как число гиперпараметров (один-два), так и время решения исходной задачи при каждом выбранном их сочетании.
+The simplest way to find the global optimum is the exhaustive search method (поиска по равномерной \cite{Bao2006} или случайной \cite{Bergstra2012} сетке).
+By selecting the number of analyzed variants for each of the hyperparameters, searching all the resulting combinations and choosing the best of them in a given metric, we obtain an estimate of the optimum with a certain accuracy \cite{Nevendra2022}.
+The exhaustive search method is easy to implement, it has perfect parallelism, but it has a significant drawback (an exponential growth of the number of grid nodes as the number of hyperparameters increases).
+Thus, exhaustive search can be used as a method for solving HPO problems only in "simple" problems where both the number of hyperparameters (one or two) and the time for solving the initial problem with each selected combination of hyperparameters are small.
 
-Альтернативой полному перебору при увеличении числа гиперпараметров являются метаэвристические (генетические и аналогичные) алгоритмы \cite{Opara2019}.
-Данные алгоритмы широко применяются при отсутствии формульного описания исследуемой модели (модель вида «черный ящик»), что характерно для рассматриваемых HPO problems \cite{Zhou2021,Yang2022}.
-Метаэвристические алгоритмы слабо зависят от числа гиперпараметров, используют информацию с предыдущих итераций для выполнения текущей, однако в силу заложенной в методы случайности дают гарантию отыскания глобального оптимума только в вероятностном смысле.
+Metaheuristic (genetic and similar) algorithms are an alternative to the exhaustive search when the number of hyperparameters increases \cite{Opara2019}.
+These algorithms are widely used in the absence of a formula-based description of the model under study (``black-box'' model), which is typical for the considered HPO problems \cite{Zhou2021,Yang2022}.
+Metaheuristic algorithms only weakly depend on the number of hyperparameters, they use information from previous iterations to perform the current one, but due to the randomness inherent in the methods, they guarantee the finding of the global optimum only in the probabilistic sense.
 
-Еще один подход к поиску глобального оптимума -- байесовская оптимизация, также применяемая для решения задач вида «черный ящик» \cite{Frazier2018,Archetti2019}.
-Фактически методы байесовской оптимизации строят стохастическую модель оптимизируемой функции.
-Модель итерационно обновляется на основе накапливаемой в процессе поиска оптимума информации, позволяя на каждой очередной итерации оценить наиболее вероятное положение глобального оптимума.
-Методы байесовской оптимизации обладают существенно более высокой эффективностью в сравнении с полным перебором, но, тем не менее, также подвержены влиянию «проклятия размерности», пусть и в меньшей степени.
+Another approach to finding the global optimum is Bayesian optimization, which is also used to solve ``black-box'' problems \cite{Frazier2018,Archetti2019}.
+In fact, Bayesian optimization methods build a stochastic model of the function being optimized.
+The model is iteratively updated on the basis of the information accumulated in the process of searching for the optimum, making it possible to estimate the most probable position of the global optimum at each iteration.
+Bayesian optimization methods have significantly higher efficiency compared to the exhaustive search, but they are also affected by the ``curse of dimensionality'', albeit to a lesser extent.
 
-В известных фреймворках по настройке гиперпараметров широко используется алгоритм Tree-Structured Parzen Estimator (TPE) \cite{hyperopt,NIPS2011}.
-TPE -- это итерационный алгоритм, идейно близкий к методам байесовской оптимизации.
-Перед началом работы алгоритма должна быть задана область поиска в пространстве гиперпараметров и определена оптимизируемая функция.
-В процессе своей работы метод проводит испытания в области поиска, после чего на их основе строит оценки распределения лучших и всех остальных результатов в пространстве гиперпараметров.
-Далее проводится поиск сочетания гиперпараметров, которое дает максимальное ожидаемое улучшение.
+In well-known hyperparameter tuning frameworks, the Tree-Structured Parzen Estimator (TPE) algorithm is widely used \cite{hyperopt,NIPS2011}.
+TPE is an iterative algorithm, conceptually close to Bayesian optimization methods.
+Before the algorithm starts, a search region in the hyperparameter space must be specified and the function to be optimized must be defined.
+In the course of its work, the method performs trials in the search region and then uses these trials to construct estimates of the distribution of the best and all other results in the hyperparameter space.
+Next, the combination of hyperparameters that gives the maximum expected improvement is searched for.
 These hyperparameters are then evaluated on the objective function.
-Описанный процесс повторяется до достижения заданного числа итераций или до тех пор, пока не будет достигнуто ограничение по времени.
+The described process is repeated until a specified number of iterations is attained or until the time constraint is reached.
 
-Хорошие результаты, достигаемые методами байесовской оптимизации  \cite{Joy2020} и TPE \cite{Watanabe2022a,Watanabe2022b}, основаны на априорном предположении, что функция соответствует определенной стохастической модели, например, Gaussian process \cite{Rasmussen2005}.
-Однако в глобальной оптимизации существуют и иные предположения о виде функции, которые дают замечательные результаты.
+The good results achieved by Bayesian optimization \cite{Joy2020} and TPE \cite{Watanabe2022a,Watanabe2022b} methods are based on the a priori assumption that the function corresponds to a certain stochastic model, such as a Gaussian process \cite{Rasmussen2005}.
+However, there are other assumptions about the shape of the function in global optimization that yield remarkable results.
 
-Одним из таких допущений о решаемой задаче является предположение об ограниченности относительных изменений целевой функции.
-В этом случае говорят, что функция удовлетворяет условию Липшица, а решаемая задача называется задачей липшицевой глобальной оптимизации.
-Для класса задач липшицевой глобальной оптимизации разработан целый ряд эффективных алгоритмов \cite{Jones2021,Paulavicius2020,Strongin2020,Sergeyev2017,PaulaviciusZilinskas2014}, которые превосходят многие другие методы глобальной оптимизации \cite{Sergeyev2018} и показали свою применимость ряде приложений \cite{Kvasov2008,CANDELIERI2019,Gubaydullin2021}.
+One of such assumptions about the problem to be solved is the assumption of boundedness of relative changes in the objective function.
+In this case, the function is said to satisfy the Lipschitz condition, and the problem to be solved is called a Lipschitz global optimization problem.
+For the class of Lipschitz global optimization problems, a number of efficient algorithms have been developed \cite{Jones2021,Paulavicius2020,Strongin2020,Sergeyev2017,PaulaviciusZilinskas2014}, which surpass many other global optimization methods \cite{Sergeyev2018} and have shown their applicability in a number of applications \cite{Kvasov2008,CANDELIERI2019,Gubaydullin2021}.
 
-Фреймворк iOpt содержит реализацию алгоритмов липшицевой оптимизации, являющихся развитием информационно-статистического алгоритма глобального поиска, детально описанного в \cite{Strongin2000,Sergeyev2013}.
-В частности, методы были адаптированы для решения задач с частично целочисленными параметрами, которые весьма часто встречаются в задачах HPO.
-Результаты экспериментов, приведенные в Section \ref{sec_exp}, показывают, что в задачах HPO фреймворк iOpt не уступает известным фреймворкам аналогичного назначения (таким как Optuna \cite{optuna} and HyperOpt \cite{hyperopt}).
+The iOpt framework contains an implementation of Lipschitz optimization algorithms, which are an extension of the information-statistical global search algorithm described in detail in \cite{Strongin2000,Sergeyev2013}.
+In particular, the methods have been adapted to solve problems with partially integer parameters, which are very common in HPO problems.
+The experimental results presented in Section \ref{sec_exp} show that the iOpt framework is not inferior in HPO problems to well-known frameworks having the same purpose (such as Optuna \cite{optuna} and HyperOpt \cite{hyperopt}).
 
@@
 
-Мы будем рассматривать задачу поиска глобального минимума в следующей постановке:
+We will consider the problem of finding the global minimum formulated as follows:
 \begin{gather}
@@
 
-Будем также предполагать, что целевая функция $\varphi(y)$ удовлетворяет в области поиска $D$ условию Липшица, т.е.
+We will also assume that the objective function $\varphi(y)$ satisfies the Lipschitz condition in the search domain $D$, i.e.
 \begin{equation} \label{f_lip}
@@
 \end{equation}
-Здесь $ \left\| \cdot \right\|$ обозначает евклидову норму, а константа $L$ является априори не известной и подлежит оценке в процессе решения. 
-%Такие задачи называются задачами липшицевой глобальной оптимизации with box-constraints.
+Here $ \left\| \cdot \right\|$ denotes the Euclidean norm, and the constant $L$ is not known a priori and is subject to estimation in the process of solution.
 
-Функция $\varphi(y)$ предполагается многоэкстремальной и заданной в виде ``черного ящика'', т.е.
-некоторого алгоритма вычисления ее значений.
-Кроме того, подразумевается, что каждое поисковое испытание (т.е.
-вычисление значения функции в точке допустимой области) требует значительных вычислительных ресурсов.
-Такая постановка проблемы полностью соответствует задаче настройки гиперпараметров методов машинного обучения \cite{Joy2020,Wang2021}.
+The function $\varphi(y)$ is assumed to be multiextremal and specified in the form of a ``black box'', i.e., of some algorithm for computing its values.
+Moreover, it is assumed that each search trial (i.e., computing the value of the function at a point in the admissible region) requires significant computational resources.
+This problem formulation fully corresponds to the task of tuning the hyperparameters of machine learning methods \cite{Joy2020,Wang2021}.
 
-Условие Липшица допускает следующую геометрическую интерпретацию.
-Допустим, что одномерная липшицева функция $f(x)$ (с известной константой Липшица $L$) была вычислена в двух точках $x'$ и $x''$.
-В соответствии с условием (\ref{f_lip}) выполняются следующие неравенства, характеризующие поведение функции $f(x)$ на интервале $[x', x'']$:
+The Lipschitz condition allows the following geometric interpretation.
+Suppose that a one-dimensional Lipschitz function $f(x)$ (with the known Lipschitz constant $L$) has been computed at two points $x'$ and $x''$.
+According to condition (\ref{f_lip}), the following inequalities characterizing the behaviour of the function $f(x)$ on the interval $[x', x'']$  are satisfied:
 \begin{gather*}
@@
 
-В силу данных неравенств значения функции в точках интервала $[x', x'']$ должны располагаться внутри области, ограниченной прямыми, проходящими через точки $(x', f(x'))$ и $(x'', f(x''))$ с наклоном $L$ и $-L$ (see Fig.
+  By virtue of these inequalities, the values of the function at the points of the interval $[x', x'']$ must lie inside the region bounded by the straight lines passing through the points $(x', f(x'))$ and $(x'', f(x''))$ with slopes $L$ and $-L$ (see Fig.
 \ref{fig1}).
@@
 \includegraphics[width=0.8\textwidth]{Fig1.pdf}
-\caption{Значения функции $f(x)$ должны находиться внутри области, выделенной цветом} \label{fig1}
+\caption{The values of the function $f(x)$ must be inside the area highlighted by colour} \label{fig1}
 \end{figure}
 
-Также в соответствии с (\ref{f_lip}) может быть записано неравенство $f(x) \geq F(x)$, где функция $F(x)$ (называемая минорантой) определяется в соответствии с формулой
+Also according to (\ref{f_lip}), we can write the inequality $f(x) \geq F(x)$, where the function  $F(x)$ (called the minorant) is defined according to the formula
+
 \[
@@
 
-Наименьшее значение миноранты $F(x)$ на $[x', x'']$ совпадает с оценкой наименьшего значения функции $f(x)$ на данном интервале. Указанная оценка достигается в точке 
+The smallest value of the minorant $F(x)$ on $[x', x'']$ coincides with the estimate of the smallest value of the function $f(x)$ on this interval. This estimate is reached at the point
 \[
@@
 \] 
-и равна
+and equals
 \[
@@
 
-Методы липшицевой глобальной оптимизации используют эту идею в своих вычислительных правилах; на ней также основано и доказательство их сходимости (see, e.g.,
+Lipschitz global optimization methods use this idea in their computational rules; the proof of their convergence is also based on this idea (see, e.g.,
 \cite{Jones2021,PaulaviciusZilinskas2014,Sergeyev2013,Evtushenko2013}).
 
-\subsection{Проблема редукции размерности}
+\subsection{The problem of dimensionality reduction}
 
-Одной из основных трудностей при решении многомерных задач глобальной оптимизации является рост вычислительных затрат при увеличении размерности задачи.
-Эта сложность имеет место и при решении задач липшицевой глобальной оптимизации.
-Например, решение задачи (\ref{f_func}) c помощью поиска по равномерной сетке с шагом $\epsilon$ потребует проведения 
+One of the main difficulties in solving multidimensional global optimization problems is the increase in computational cost as the dimensionality of the problem increases.
+This difficulty also occurs when solving Lipschitzian global optimization problems.
+For example, solving problem (\ref{f_func}) by means of the uniform grid search with a step $\epsilon$ will require performing
 \[
@@
 \]
-испытаний. Очевидно, что $K$ экспоненциально возрастает с ростом $N$.
-Уменьшение числа испытаний при тех же требованиях к точности решения возможно за счет адаптивного построения неравномерной сетки в области изменения параметров.
+trials.
+Obviously, $K$ increases exponentially with increasing $N$.
+It is possible to reduce the number of trials with the same requirements to the solution accuracy using the adaptive construction of a non-uniform grid in the region of parameter variation.
 
-Известным подходом к решению многомерной задачи (\ref{f_func}) является ее редукция к одномерной задаче с последующим применением эффективных одномерных алгоритмов глобальной оптимизации (see \cite{Strongin2000,Sergeyev2013}).
-Здесь хорошо зарекомендовал себя подход, в котором размерность задачи редуцируется при помощи кривой Пеано–Гильберта $y(x), x \in [0, 1]$.
-Такая кривая однозначно и непрерывно отображает интервал $[0, 1]$ на гиперинтервал $D$ из (\ref{f_D}), т.е. заполняет всю область $D$.
+A well-known approach to solving the multidimensional problem (\ref{f_func}) is to reduce it to a one-dimensional problem and then apply efficient one-dimensional global optimization algorithms (see \cite{Strongin2000,Sergeyev2013}).
+A well-proven approach here is to reduce the dimensionality of the problem by means of a Peano-Hilbert curve $y(x), x \in [0, 1]$.
+Such a curve uniquely and continuously maps the interval $[0, 1]$ onto the hyperinterval $D$ of (\ref{f_D}), i.e., it fills the entire region $D$.
 
-Для решения задачи глобальной липшицевой оптимизации кривые, заполняющие пространство, могут быть применены следующим образом.
-Если $y(x)$ есть кривая Пеано–Гильберта, то из непрерывности целевой функции $\varphi(y)$ следует, что
+To solve the Lipschitz global optimization problem, space-filling curves can be applied as follows.
+If $y(x)$ is a Peano-Hilbert curve, then it follows from the continuity of the objective function $\varphi(y)$ that
 \[
@@
 \]
-т.е. исходная многомерная задача (\ref{f_func}) редуцируется к одномерной.
+i.e., the original multidimensional problem (\ref{f_func}) is reduced to a one-dimensional problem.
 
-При этом известно (see \cite{Strongin2000}), что если многомерная функция $\varphi(y), \; y \in D,$  удовлетворяет условию Липшица с константой $L$, то редуцированная одномерная функция $\varphi(y(x)), \; \in [0,1]$ удовлетворяет the H{\"o}lder condition
+It is known (see \cite{Strongin2000}) that if a multidimensional function  $\varphi(y), \; y \in D$,  satisfies the Lipschitz condition with the constant $L$, then the reduced one-dimensional function $\varphi(y(x)), \; x \in [0,1]$ satisfies the H{\"o}lder condition
 \[
@@
 \]
-Здесь $N$ есть размерность исходной задачи, а коэффициент $ H=2 L \sqrt{N+3}$.
+Here, $N$ is the dimensionality of the original problem, and the coefficient
+$ H=2 L \sqrt{N+3}$.
 
-Замечание 1. Для редуцированной одномерной функции $\varphi(y(x))$  не будет выполняться условие Липшица. Однако многие одномерные алгоритмы липшицевой оптимизации
-могут быть обобщены на случай минимизации гeльдеровых функций. Конкретный пример такого обобщения приведен в subsection \ref{sec_GSA}.
+Remark 1.
+The Lipshitz condition will not be satisfied for the reduced one-dimensional function $\varphi(y(x))$.
+However, many one-dimensional Lipschitz optimization algorithms can be generalized to the case of H{\"o}lder function minimization.
+A concrete example of such a generalization is given in subsection \ref{sec_GSA}.
 
@@
 
-Замечание 2.
-Теоретическая кривая Пеано-Гильберта $y(x)$ определяется как предельный объект.
-В численных алгоритмах применяются evolvents $y_m(x)$, аппроксимирующие истинную curve c заданным уровнем плотности $2^{-m}$, зависящим от требуемой точности поиска.
-Эффективные схемы вычисления нескольких видов evolvents подробно описаны в \cite{Sergeyev2013}.
-Для иллюстрации на Fig.~\ref{fig:Peano} приведены графики, соответствующие разверткам $y_m(x)$ с плотностью $m=4$ для размерностей $N=2$ и $N=3$.
+Remark 2: The theoretical Peano-Hilbert curve $y(x)$ is defined as a limiting object.
+Numerical algorithms use evolvents $y_m(x)$, that approximate the true curve with a given density level $2^{-m}$, which depends on the required search accuracy.
+Efficient schemes for computing several types of evolvents are described in detail in \cite{Sergeyev2013}.
+For illustration, Fig.~\ref{fig:Peano} shows graphs corresponding to the evolvents $y_m(x)$ with density $m=4$ for dimensions $N=2$ and $N=3$.
 
-\subsection{Проблема дискретных параметров}
+\subsection{The problem of discrete parameters}
 \label{sec_discr}
 
-Особый интерес представляют задачи, в которых часть параметров может принимать значения из заранее заданного конечного множества, т.к.
-для них сложнее построить оценки оптимума по сравнению с задачами с непрерывными параметрами.
-Методам решения задач со смешанными параметрами посвящено много публикаций (см., например, обзоры \cite{Burer2012,Boukouvala2016}). 
-Известные детерминированные методы решения задач данного класса ориентированы на решение линейных или выпуклых задач \cite{Lee2012}.
-Сложные многоэкстремальные задачи предлагается решать с помощью эвристических подходов \cite{Belotti2013}.
+Of particular interest are problems where some of the parameters can assume values from a predetermined finite set, since it is more difficult to construct estimates of the optimum for these problems than for problems with continuous parameters.
+Many publications address the methods for solving mixed-parameter problems (see, for example, reviews \cite{Burer2012,Boukouvala2016}).
+Known deterministic methods for solving problems of this class are centered on solving linear or convex problems \cite{Lee2012}.
+It is suggested that complex multi-extremal problems should be solved using heuristic approaches \cite{Belotti2013}.
 
-Нами был предложен и реализован новый детерминированный подход к решению задач со смешанными параметрами; приведем здесь его краткое описание. 
-Пусть  целевая функция задачи зависит от двух векторов параметров: вектора $y$, принадлежащего гиперинтервалу $D$, и вектора $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е.
+We have proposed and implemented a new deterministic approach to solving problems with mixed parameters; here we give a brief description of it. 
+Let the objective function of the problem depend on two parameter vectors: vector $y$, which belongs to the hyperinterval $D$, and vector $u$, which has a finite (and not very large) set of possible values, i.e.
 \begin{gather}\label{problem_i}
@@
 
-Такие конечные наборы значений могут характеризовать множество возможных сочетаний категориальных гиперпараметров исследуемого алгоритма машинного обучения. 
-Характерными примером является support vector machine, влияние на качество работы которого оказывают непрерывные параметр регуляризации $C$ и коэффициент ядра $\gamma$.
-При этом тип функции ядря выбирается из конечного множества вариантов (например, полиномиальная функция, радиальная базисная функция, сигмоид и т.д.).
+Such finite sets of values can characterize many possible combinations of categorical hyperparameters of the machine learning algorithm under study.
+A typical example is the support vector machine, whose performance is influenced by the continuous regularization parameter $C$ and the kernel coefficient $\gamma$.
+In this case, the type of kernel function is chosen from a finite set of variants (e.g., polynomial function, radial basis function, sigmoid, etc.).
 %Целочисленные параметры = непрерывные параметры с округлением, добавить комментарий об этом?
 
-Занумеруем целыми числами $s, 1\leq s \leq S,$ все различные комбинации категориальных параметров, т.е. сопоставим каждому номеру $s$ вектор $u_s$. 
-Тогда рассматриваемая задача может быть записана в виде 
+Let us number all the different combinations of categorical parameters with integers $s, 1\leq s \leq S$, i.e., let us assign a vector $u_s$ to each number $s$.
+Then the problem under consideration can be written in the form
 \begin{gather}\label{problem_is}
@@
 
-Используя схему редукции размерности с помощью кривых Пеано $y(x), x\in [0,1]$ можно сопоставить каждой задаче минимизации по $y$ одномерную задачу минимизации
+Using a dimensionality reduction scheme with Peano curves $y(x), x\in [0,1]$,  we can assign to each minimization problem on $y$ a one-dimensional minimization problem
 \[
@@
 
-Рассмотрим теперь отображение 
+Let us now consider the mapping
 \[
@@
 \]
-переводящее любую точку интервала [0,S] на область $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функцию 
+which translates any point of the interval $[0,S]$ to the region $D$ (the notation $E(x)$ corresponds to the integer part of the number $x$) and define the function
 \[
@@
 \]
-имеющую, вообще говоря, разрывы в целочисленных точках $x_i = i, 1\leq i \leq S-1$.
-Поэтому значения  $z_i = f(x_i)$ в этих точках будут считаться неопределенными и не будут использоваться в алгоритме.
+which has, generally speaking, discontinuities at integer points $x_i = i, 1 \leq i \leq S-1$.
+Therefore, the values $z_i = f(x_i)$ at these points will be considered undefined and will not be used in the algorithm.
 %, а значения индекса -- равным 0, т.е. $\nu(x_i) = 0$.
 
-Используя введенные обозначения можно переформулировать исходную задачу как
+Using the introduced notations, we can reformulate the original problem as
 \begin{equation}\label{problem_is1}
@@
 
-Применяя к решению задачи (\ref{problem_is1}) алгоритм глобального поиска, мы найдем решение задачи (\ref{problem_i}).
-При этом основная часть испытаний будет проведена в $s$-й подзадаче, решение которой соответствует решению исходной задачи (\ref{problem_i}).
-В остальных подзадачах будет проведена лишь незначительная часть испытаний, т.к.
-решения данных подзадач являются локально-оптимальными по отношению к решению $s$-й подзадачи.
+Applying the global search algorithm to the solution of problem (\ref{problem_is1}), we will find the solution of problem (\ref{problem_i}).
+In this case, the main part of trials will be carried out in the $s$-th subproblem, whose solution corresponds to the solution of the original problem (\ref{problem_i}).
+In the other subproblems, only a small part of trials will be carried out, since the solutions of these subproblems are locally optimal with respect to the solution of the $s$-th subproblem.
 
@@
 
-Фреймворк методов интеллектуальной оптимизации iOpt разработан в рамках выполенния проекта ``Strong Artificial Intelligence in Industry''.
-Он предназначен для выбора оптимальных (в заданной метрике) значений параметров сложных объектов и процессов, например, методов искусственного интеллекта и машинного обучения.
-Фреймворк позволяет проводить точную настройку параметров моделей и методов, используемых в прикладных исследованиях в различных научных областях.
-Фреймворк iOpt разработан на языке программирования Python с использованием его стандартной библиотеки.
-Для его использования необходим установленный интерпретатор языка Pyhon версии не ниже 3.8.
-Фреймворк является open source и доступен по адресу https://github.com/aimclub/iOpt.
+The iOpt framework of intelligent optimization methods was developed as part of the ``Strong Artificial Intelligence in Industry'' project.
+It is designed to select optimal (in a given metric) values of parameters of complex objects and processes, for example, artificial intelligence and machine learning methods.
+The framework allows fine-tuning of parameters of models and methods used in applied research in various fields of science.
+The iOpt framework was designed in the Python programming language using its standard library.
+To use the framework, you need to have the Pyhon interpreter version 3.8 or higher installed.
+The framework is open source and is available at \url{https://github.com/aimclub/iOpt}.
 
-Фреймворк iOpt может использоваться при создании специализированных систем поддержки принятия решений для решения задач выбора оптимальных параметров сложных моделей или методов.
-Область применения фреймворка фокусируется на задачах настройки гиперпараметров, поскольку при решении задач этого класса возникают проблемы многоэкстремальности критерия оптимизации и отсутствия формульного описания исследуемой модели (модель вида «черный ящик»).
+The iOpt framework can be used to develop specialized decision support systems for solving the tasks of selecting optimal parameters for complex models or methods.
+The application area of the framework is focused on the hyperparameter tuning tasks, since this class of tasks involves the problems of multiextremality of the optimization criterion and the lack of a formula-based description of the model under study (a ``black box'' type model).
 
@@
 
-Алгоримтическим ядром фреймворка iOpt является информационно-статистический алгоритм глобального поиска для решения задач вида (\ref{f_func}), (\ref{f_D}). 
-Данный алгоритм предполагает построение последовательности точек $y^i$, в которых проводятся поисковые испытания, т.е.
-вычисляются значения objective function $z^i = \varphi(y^i)$.
-В соответствии с используемой схемой редукции размерности проведение испытания подразумевает вычисление $y^i=y(x^i))$, поэтому результатом испытания будет являться набор значений $(x^i, y^i=y(x^i), z^i = \varphi(y^i))$.
-Поисковую информацию, накопленную после проведения $k$ испытаний, удобно представлять в виде множества, в котором соответствующие наборы перенумерованы (нижним индексом) в порядке возрастания $x$, т.е.
-в виде множества
+The algorithmic core of the iOpt framework is an information-statistical global search algorithm for solving problems of the form (\ref{f_func}), (\ref{f_D}).
+This algorithm assumes the construction of a sequence of points $y^i$,  at which search trials are performed, i.e., the values of the objective function $z^i = \varphi(y^i)$ are computed.
+According to the dimensionality reduction scheme used, conducting a trial implies computing $y^i=y(x^i)$, so the result of the trial will be a set of values $(x^i, y^i=y(x^i), z^i = \varphi(y^i))$.
+It is convenient to represent the search information accumulated after conducting $k$ trials as a set in which the corresponding entries are renumbered (by the lower index) in ascending order of $x$, i.e., as a set
 \begin{equation}\label{omega}
@@
 \end{equation}
-где $x_1 < x_2 < ... < x_k$.
+where $x_1 < x_2 < ... < x_k$.
 
-Основная идея алгоритма глобального поиска состоит в следующем.
-Пусть в процессе поиска проведено $k \geq 2$ испытаний and the search information $\Omega_k$ from (\ref{omega}) is obtained.
-Для выбора точки очередного испытания для каждого поискового интервала $(x_{i-1},x_i), 1<i\leq k$, вычисляется его характеристика в соответствии с формулой
+The main idea of the global search algorithm is as follows.
+Let $k \geq 2$ trials be conducted in the search process and the search information $\Omega_k$ from (\ref{omega}) is obtained.
+To select the point of the next trial for each search interval $(x_{i-1},x_i), 1<i\leq k$, its characteristic is calculated according to the following formula
 \begin{equation}\label{R}
@@
 \end{equation}
-Здесь $r>1$ -- параметр алгоритма, $\Delta_i=(x_i-x_{i-1})^{1/N}$.
-Значение $\mu$ является нижней оценкой константы Гельдера целевой функции, вычисленной на основе поисковой информации из (\ref{omega}) по формуле
+
+Here,  $r>1$ is the algorithm parameter, $\Delta_i=(x_i-x_{i-1})^{1/N}$.
+The value $\mu$ is the lower estimate of the H\"older constant of the objective function computed using the search information from (\ref{omega}) by the formula
 \begin{equation}\label{mu}
@@
 \end{equation}
-причем в случае, если формула (\ref{mu}) выдает значение $\mu=0$, то используется значение $\mu=1$.
+and in case when formula (\ref{mu}) gives the value $\mu=0$, the value $\mu=1$ is used.
 
-Новое испытание проводится в точке $x^{k+1}$, принадлежащей интервалу с наибольшей характеристикой.
-Если правая точка этого интервала имеет номер $t$ в множестве $\Omega_k$, то формула для вычисления $x^{k+1}$ будет выглядеть следующим образом:
+A new trial is performed at the point $x^{k+1}$ that belongs to the interval
+with the largest characteristic. If the right point of this interval has the number $t$ in the set $\Omega_k$, then the formula for calculating $x^{k+1}$ will look as follows:
 \begin{equation}\label{xk1}
@@
 Algorithm \ref{alg_GSA} shows the pseudo code of the Global Search Algorithm.
-Входными данными алгоритма являются objective function $\varphi(y)$  и область поиска $D$ из (\ref{f_func}) and (\ref{f_D}), параметр алгоритма $r>1$, точность поиска минимума $\epsilon > 0$, максимально допустимое число испытаний $K_{max}$.
-Выходным параметром алгоритма является оценка точки глобального минимума $y_{min}$.
-Дополнительными выходными параметрами могут быть значение функции в точке минимума $\varphi(y_{min})$ и число проведенных испытаний $k$.
+The input data of the algorithm are the objective function $\varphi(y)$ and the search region $D$ from (\ref{f_func}) and (\ref{f_D}), the algorithm parameter $r>1$, the minimum search accuracy $\epsilon > 0$, and the maximum allowable number of trials $K_{max}$.
+The output parameter of the algorithm is an estimate of the global minimum point $y_{min}$.
+Additional output parameters can include the value of the function at the minimum point $\varphi(y_{min})$ and the number of trials $k$.
 
@@
  \While{ $\Delta_t \geq \epsilon$ and  $k \leq K_{max}$}{
-  Вычислить оценку H\"older constant $\mu$ в соответствии с (\ref{mu})\\
+  Compute the estimate of the H\"older constant $\mu$ according to (\ref{mu})\\
 	\For{$i=1$ \KwTo $k$}{
-	Вычислить характеристику $R(i)$ в соответствии с (\ref{R})\\
+	Compute the characteristic $R(i)$ according to (\ref{R})\\
 	}
-	Определить номер $t$, для которого $R(t) = \max \left\{ R(i), 1 \leq i \leq k \right\}$
+	Determine the number $t$, for which $R(t) = \max \left\{ R(i), 1 \leq i \leq k \right\}$
 
-	Вычислить точку нового испытания $x^{k+1}$ в соответствии с (\ref{xk1})
+	Compute the new trial point $x^{k+1}$ according to (\ref{xk1})
 
-	Провести новое испытание в точке $x^{k+1}$
+	Perform a new trial at the point $x^{k+1}$
 
-	Добавить $(x^{k+1}, y^{k+1}, z^{k+1})$ в множество $\Omega_k$
+	Add $(x^{k+1}, y^{k+1}, z^{k+1})$ to the set $\Omega_k$
 
@@
  \KwRet{$y_{min}$}
- \caption{Global search}\label{alg_GSA}
+ \caption{Global search algorithm}\label{alg_GSA}
 \end{algorithm}
 
-Теория сходимости алгоритма, а также некоторые его модификации представлены в \cite{Strongin2000}.
+The theory of algorithm convergence as well as some of its modifications are presented in \cite{Strongin2000}.
 
@@
 
-Рассмотрим случай, когда в настраиваемом алгоритме имеются как непрерывные, так и категориальные параметрами. 
-В этом случае категориальным параметрам можно поставить в соответствие целочисленные значения, и свести сзадачу к постановке (\ref{problem_is1}).
-Так как решение задач с частично целочисленными параметрами в соответсвии со схемой, изложенной в Section \ref{sec_discr}, предполагает решение задачи (\ref{problem_is1}) с разрывной целевой функцией, требуется модификация алгоритма, учитывающая эту особенность.
+Let us consider the case when the algorithm to be tuned has both continuous and categorical parameters.
+In this case, the categorical parameters can be assigned integer values, and the problem can be reduced to the formulation (\ref{problem_is1}).
+Since the solution of problems with partially integer parameters in accordance with the scheme described in Section \ref{sec_discr} implies the solution of problem (\ref{problem_is1}) with a discontinuous objective function, a modification of the algorithm that takes this feature into account is required.
 
-С этой целью введем классификацию точек из области поиска с помощью специального индекса $\nu_i=\nu(x_i)$.
-Индексы целочисленных точек $x_i = i, 0\leq i \leq S$, будем полагать равным нулю, т.е.
+For this purpose, we introduce the classification of points from the search region using a special index $\nu_i=\nu(x_i)$.
+We will assume the indices of integer points $x_i = i, 0\leq i \leq S$, to be equal to zero, i.e.
 $\nu(x_i) = 0$.
-Для всех остальных значений $x\in(i,i+1),  0 \leq i < S$, значение индекса будем считать равным единице, т.е.
+For all other values of $x\in(i,i+1),  0 \leq i < S$, we will assume the index value to be equal to unity, i.e.
 $\nu(x) = 1$.
-Тогда результатом испытания в некоторой точке $x^j\in[0,S]$ будет являться набор значений
-$(x^j, \nu^j=\nu(x^j), y^j=y(x^j), z^j = \varphi(y^j))$. Причем в случае $\nu^j=0$ значение $z^j$ будет считаться неопределенным.
+Then the result of the trial at some point $x^j\in[0,S]$  will be the set of values $(x^j, \nu^j=\nu(x^j), y^j=y(x^j), z^j = \varphi(y^j))$.
+In the case of  $\nu^j=0$, the value of $z^j$ will be considered undefined.
 
-Общая вычислительная схема, изложенная в виде Алгоритма \ref{alg_GSA}, останется прежней. Модификации будут состоять в следующем.
-Модификации будут состоять в следующем.
+The general computational scheme outlined as Algorithm \ref{alg_GSA} will remain the same. The modifications will consist of the following.
 
 1.
-На этапе инициализации в множество $\Omega_k$, содержащее поисковую информацию, заносятся значения $(x^i = i, \nu^i=0, y^i=y(x^i), z^i)$, $ i, 0\leq i \leq S$, т.е.
-заносятся граничные точки интервалов $(x_{i-1},x_i), 1<i\leq S$.
-После этого проводятся испытания в произвольных внутренних точках (например, в серединах) интервалов $(x_{i-1},x_i), 1<i\leq S$; результаты испытаний также добавляются в поисковую информацию. 
-Таким образом, после инициализации множество $\Omega_k$ будет содержать $k=2S+1$ элемент.
+At the initialization stage, the values $(x^i = i, \nu^i=0, y^i=y(x^i), z^i)$, $ i, 0\leq i \leq S$, i.e., the boundary points of the intervals $(x_{i-1},x_i), 1<i\leq S$, are entered into the set $\Omega_k$ containing the search  information.
+After that, trials are carried out at arbitrary internal points (e.g., midpoints) of the intervals $(x_{i-1},x_i), 1<i\leq S$; the trial results are also added to the search information.
+Thus, after initialization, the set $\Omega_k$ will contain $k=2S+1$ elements.
 
 2.
-Вычисление оценки константы Гельдера $\mu$ в соответствии с формулой (\ref{mu}) должно выполняться только для  интервалов, для которых $\nu(x_i) = \nu(x_{i-1}) = 1$, т.е.
-когда значения $z_i$ and $z_{i-1}$ определены.
+The computation of the estimate of the H\"older constant $\mu$ according to formula (\ref{mu}) must be performed only for intervals for which $\nu(x_i) = \nu(x_{i-1}) = 1$, i.e., when the values $z_i$ and $z_{i-1}$ are determined.
 
-3. Вычисление характеристики поискового интервала $(x_{i-1},x_i), 1<i\leq k$, проводится по следующим правилам:
+3. The computation of the characteristic of the search interval $(x_{i-1},x_i), 1<i\leq k$, is performed according to the following rules:
 \begin{gather}\label{R_int}
@@
 \end{gather}
-Отметим, что в процессе работы алгоритма не будет возникать интервалов, у которых обе граничные точки имеют неопределенные значения функции, т.е. $\nu(x_i) = \nu(x_{i-1}) = 0$.
 
-4. Точка нового испытания $x^{k+1}$ будет вычисляться следующим образом:
+Note that the algorithm will not produce intervals whose both boundary points have undefined function values, i.e., $\nu(x_i) = \nu(x_{i-1}) = 0$.
+
+4. The new trial point $x^{k+1}$ will be calculated as follows:
 \begin{gather}\label{xk1_int}
@@
 \end{gather}
-Здесь $t$ есть номер поискового интервала, имеющего максимальную характеристику.
 
-Еще раз отметим, что характеристика $R(i)$ количественно оценивает возможность нахождения точки $x^*$, являющейся inverse image of the global optimizer $y* = y(x^*)$, в пределах рассматриваемого интервала $(x_{i-1},x_i)$.
+Here, $t$ is the number of the search interval that has the maximum characteristic.
+
+We note again that the characteristic $R(i)$ quantifies the possibility of finding a point $x^*$, which is the inverse image of the global optimizer $y^* = y(x^*)$, within the considered interval $(x_{i-1},x_i)$.
 

### 2023-08-29 2023/MMST/samplepaper.tex
change the section 3 header

@@
 
-\section{Solving Method}
+\section{Method for Solving the Problems with Partially Defined Objective Function}
 
@@
 \end{equation}
-        \paragraph{Case 2.}   For the intervals with \textbf{one non-computable point and one boundary point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
+        \paragraph{Case 2.}   For the intervals with \textbf{one non-computable and one boundary point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
 \begin{equation}\label{eq22} 

### 2023-08-30 2023/ITMO paper/elsarticle-template-num.tex
ru ->en path2

@@
 \subsection{Experimental setup}
+In this section, we present experimental results that demonstrate the effectiveness of the developed framework for a number of HPO problems.
+First, we describe in detail the setup of the computational experiment.
+Then, the results of comparing the iOpt framework with several well-known frameworks widely used for hyperparameter tuning in machine learning methods are presented.
+The section closes with conclusions and a discussion of the results.
 
-Datasets
+\textbf{Datasets.} 
+We used for our experiments fifteen datasets from the UCI repository \cite{AsuncionNewman2007}, which covers a large number of areas.
+Datasets from this repository are quite often used to test the correct operation of hyperparameter tuning algorithms \cite{WU2023110160,Wang2021,Joy2020}.
+We selected the datasets to ensure that the data contained in them were described by both numerical and categorical attributes.
+Note that most of the datasets were unbalanced and the data presented in the datasets were adapted for the classification task.
+The characteristics of the datasets used are presented in Table \ref{tab:1}.
 
-\begin{table}[]
-\caption{14 datasets used in the experiments.}
+\begin{table}[ht]
+\caption{The datasets used in the experiments}
 \label{tab:1}
@@
 
-Metrics
+\textbf{Classification algorithms.} We performed hyperparameter tuning for three types of classification algorithms on all datasets.
+\begin{enumerate}
+	\item XGBoost is one of the tree ensemble construction methods, which is an efficient implementation of gradient boosting over decision trees \cite{XGBoost2016}.
+The main advantages of the algorithm include a smart penalty mechanism, proportional reduction of leaf nodes, and automatic feature selection.
+It is known that the performance of an algorithm and its predictive ability depend significantly on the correct tuning of hyperparameters.
+The implementation of the method is based on the widely used XGBoost library.
+ \item Multi-Layer Perceptron (MLP), a fully connected artificial neural network, is a supervised learning algorithm that forms a functionf:$R^m \to R^o$, where $m$ is the number of input attributes, and o is the output dimensionality of the network \cite{GARDNER19982627}.
+Using this method, it is possible to form the function f as a universal nonlinear approximator for the classification and regression problem.
+In the framework of this study, we will consider the multilayer perceptron for the task of attributing an object to a particular class.
+ \item Support Vector Classification (SVC) is a support vector method for the classification problem \cite{SVC2003}.
+The main purpose of this algorithm is to find the equation of a separating hyperplane $(\sum_{i=1}^n \omega_i x_i)+\omega_0=0$ in the space $R^n$ that separates classes in some optimal way.
+To improve the predictive ability of the algorithm, a method for constructing a nonlinear classifier was proposed by switching from scalar products to arbitrary kernels, which allows the construction of nonlinear separators.
+An implementation of the SVC method from the scikit learn library was used to investigate the capabilities of HPO algorithms.
 
-% Please add the following required packages to your document preamble:
-% \usepackage{multirow}
-\begin{table}[]
-\caption{Настраиваемые гиперпараметры для алгоритмов машинного обучения}
+\end{enumerate}
+
+The tunable hyperparameters of the above algorithms and their search region are listed in Table \ref{tab:2}.
+
+\begin{table}[ht]
+\caption{Tunable hyperparameters of machine learning algorithms.}
 \label{tab:2}
@@
 \hline
-Метод                    & Параметры            & Тип параметра & Диапазон параметра                                                     \\ \hline
+Method  & Parameters & Parameter type & Parameter range \\ \hline
 \multirow{7}{*}{XGBoost} & n\_estimators        & Int           & {[}10, 200{]}                                                          \\
@@
 
-Algorithms for comparison
+\textbf{Evaluation Metrics.} To evaluate the quality of machine learning methods on the considered datasets, it is necessary to determine the metrics.
+In view of the peculiarities of datasets, namely their unbalanced nature and the different number of classes in different datasets, the use of a single metric would result in a biased evaluation.
+Therefore, we used the following set of metrics:
+\begin{itemize}
+	\item Accuracy is a metric describing the overall prediction accuracy of the method across all classes.
+It is most visible in the case when each class is equally important and the dataset shows a balance of classes \cite{Metric1}.
+This metric has been applied when working with the datasets Optical Recognition of Handwritten Digits, Bank Marketing, CNAE9, Statlog Segmentation, Semeion Handwritten Digit;
+ \item Binary F1 - F1 score is adapted for binary classification tasks, with the predominance of one class over another in the dataset \cite{Chicco2020}.
+This metric is suitable for analyzing Breast Cancer Wisconsin, Credit Approval, Parkinsons, Banknote authentication, Wilt datasets;
+ \item Macro F1 - F1 score is suitable for tasks with a large number of unbalanced classes \cite{Metric2}.
+The datasets Protein Localization Sites (Ecoli), Balance Scale Weight \& Distance, Zoo, Car Evaluation, Turkiye Student Evaluation have this particular feature.
 
-\subsection{Experimental results}
+\end{itemize}
 
-First result
+\textbf{HPO algorithms.} The hyperparameters of machine learning methods were tuned using the iOpt framework developed by us, and also using the well-known frameworks Optuna \cite{optuna} and HyperOpt \cite{hyperopt}.
 
-Second result
+Optuna is an automatic hyperparameter optimization software environment specifically designed for optimizing hyperparameters of machine learning algorithms.
+Several methods for selecting test points, which are referred to as \textit{samplers} within this framework, are available in Optuna.
+In the study, we used the following hyperparameter search options:
+\begin{enumerate}
+	\item Optuna(random) will perform a random search. For each hyperparameter, a distribution is specified, from which its value is selected.
+	\item Optuna(tpe) is a sampler based on the Tree-structured Parzen Estimator algorithm \cite{NIPS2011_86e8f7ab}.
+On each trial, for each parameter, TPE fits one Gaussian Mixture Model (GMM) $l(x)$ to the set of parameter values associated with the best objective values, and another GMM $g(x)$ to the remaining parameter values.
+It chooses the parameter value x that maximizes the ratio $l(x)/g(x)$.
+	\item Optuna(cmaes) is an algorithm based on CMA-ES (Lightweight Covariance Matrix Adaptation Evolution Strategy) \cite{cmaes2015}.
+The CMA-ES is a stochastic method for real-parameter optimization of non-linear, non-convex functions.
+The algorithm derived from intuitive concepts and from requirements of non-linear, non-convex search in continuous domain.
+	\item Optuna (nsgaii) is a sampler based on the Nondominated Sorting Genetic Algorithm II,  which is a well known, fast and elitist multi-objective genetic algorithm \cite{nsgaii2013}.
 
-Third result
+\end{enumerate}
 
-\begin{table}[]
-\caption{Время поиска значений гиперпараметров при однократном запуске алгоритма оптимизации.}
+Hyperopt framework is a Python library for optimization over complex search spaces, which may include real-valued, discrete, and conditional dimensions.
+As with Optuna, Hyperopt has different options for finding hyperparameter values.
+There are currently three algorithms implemented in Hyperopt: Random Search, Tree of Parzen Estimators (TPE) and Adaptive TPE.
+We used the computation variant with the TPE algorithm implementation - Hyperopt(tpe) \cite{NIPS2011_86e8f7ab}.
+
+When using the iOpt framework, we applied a variant of computation in which the available resource was divided into portions by the number of trials.
+Most of the trials (75\%) were performed using the global search algorithm (see Section \ref{sec_iOpt}), while in the remaining trials we used the local method to refine the solution found \cite{Kelley}.
+
+\subsection{Results and discussion}
+
+Computational experiments were performed on the computing nodes of the Lobachevsky supercomputer (Intel Xeon Silver 4310T CPU @ 2.30GHz, 64Gb RAM, Linux CentOS 7).
+We used the following versions of frameworks for the experiments: iOpt 0.2.22, Optuna 3.1.1, Hyperopt 0.2.7.
+All runs were performed using Python version 3.11.3.
+
+In one experiment on hyperparameter tuning, we ran 200 search trials with each method.
+Due to the stochastic nature of the Optuna and Hyperopt algorithms, we ran them 10 times and calculated both the average and worst values of the objective  metric.
+For iOpt, we ran it once because the global search algorithm is deterministic and the average, worst and best case values are the same.
+
+Figures \ref{fig_res_1} and \ref{fig_res_2} show the values of the classification metrics of the SVC algorithm in the worst and average case, respectively.
+As can be seen, when optimizing the hyperparameters of the SVC algorithm on the Banknote, CNAE-9, Statlog Segmentation, and Wilt datasets, the iOpt framework has demonstrated the best results.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{SVC_1.pdf}
+\caption{Values of the optimized metrics for the SVC algorithm, obtained in the worst case} \label{fig_res_1}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{SVC_2_avg.pdf}
+\caption{Values of optimized metrics for the SVC algorithm averaged over 10 runs} \label{fig_res_2}
+\end{figure}
+
+On the CarEvaluation, Credit Approval, Ecoli, Parkinsons, and Semeion datasets, the best values of the metrics on average were obtained with the Optuna (tpe).
+On the datasets Balance, Breast Cancer, Digits, on average the Optuna(cmaes) performed well.
+The Hyperopt(tpe) performed better than the others on the Bank-marketing dataset.
+In the case of the Zoo dataset, all optimization algorithms showed the same results.
+
+It is especially worth noting that the algorithms implemented in both Optuna and Hyperopt contain random parameters; as a consequence, the best values of the objective metric are not guaranteed for a single run.
+
+This fact is illustrated by Fig.
+\ref{fig_res_1}.
+The iOpt framework additionally performed better on the Bank-marketing, Banknote, CNAE-9, Statlog Segmentation, and Wilt datasets.
+Thus on 9 out of 14 datasets the iOpt framework showed better values of the objective metric compared to the Optuna and Hyperopt frameworks.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{XGBoost_1.pdf}
+\caption{Values of the optimized metrics for the XGBoost algorithm, obtained in the worst case} \label{fig_res_3}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{XGBoost_2_avg.pdf}
+\caption{Values of optimized metrics for the XGBoost algorithm averaged over 10 runs} \label{fig_res_4}
+\end{figure}
+
+Figures \ref{fig_res_3} and \ref{fig_res_4} show the value of classification metrics of the XGBoost algorithm in the worst and average case, respectively.
+In the case of optimizing the hyperparameters  of the XGBoost algorithm, the iOpt framework showed better results on the CarEvaluation, CNAE-9, Credit Approval, Ecoli, Parkinsons, Wilt, Zoo datasets.
+
+On average, when using 10 runs, the best performance of the Optuna(tpe) framework was observed on the Balance, Bank-marketing, Banknote, Breast Cancer, CarEvaluation, Credit Approval, Digits, Ecoli, Semeion, Statlog Segmentation, Wilt datasets.
+The Hyperopt(tpe) showed good results on the CNAE-9 and Parkinsons datasets on average for 10 runs.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{MLP_1.pdf}
+\caption{Values of the optimized metrics for the XGBoost algorithm, obtained in the worst case} \label{fig_res_5}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{MLP_2_avg.pdf}
+\caption{Values of optimized metrics for the XGBoost algorithm averaged over 10 runs} \label{fig_res_6}
+\end{figure}
+
+Figures \ref{fig_res_5} and \ref{fig_res_6} show the values of classification metrics of the MLP algorithm.
+The iOpt framework performed better on the Balance, Bank-marketing, Banknote, Breast Cancer, Credit Approval, Wilt, Zoo datasets.
+Considering the average values of metrics on 10 runs of hyperparameter optimization, the Optuna framework performed better on the Breast Cancer, CNAE-9, Credit Approval, Digits, Ecoli, Parkinsons, Semeion datasets, and the Hyperopt on the Statlog Segmentation, Wilt, CarEvaluation datasets.
+
+Thus, we can conclude that the iOpt framework shows comparable results with the Optuna and Hyperopt frameworks in terms of performance quality.
+It should be reminded that the algorithms implemented within the iOpt framework are deterministic.
+As a consequence, iOpt, unlike Optuna and Hyperopt, only needs to be run once to achieve the best values of the quality metric.
+
+Next, we consider the performance of the iOpt, Optuna, and Hyperopt frameworks.
+Table \ref{tab:3} shows the running time of the frameworks for optimizing the hyperparameters of the SVC algorithm.
+A single optimization run was performed with a maximum number of 200 search trials.
+It is obvious that even in the case of a single run iOpt shows better performance on many datasets compared to Optuna and Hyperopt.
+
+To obtain a good solution, the randomized methods from Optuna and Hyperopt should be repeated several times.
+In this case, the performance gain of iOpt is even greater.
+As an example, Figure \ref{fig8} shows a comparison of the runtime of the frameworks with 10 reruns of Optuna and Hyperopt on the Wilt dataset.
+
+\begin{table}[ht]
+\caption{Time of hyperparameter tuning for a single run of algorithms.}
 \label{tab:3}
@@
 \end{table}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{Fig8}
+\caption{Time for tuning hyperparameters of the SVC algorithm to obtain good values of quality metrics on the Wilt dataset} \label{fig8}
+\end{figure}
+
+\section{Conclusion}
+In this paper, we propose a hyperparameter tuning method that is based on the ideas of Lipschitz global optimization.
+In particular, we use a global search algorithm combined with dimensionality reduction based on Peano curves.
+We have proposed a generalization of Lipschitz optimization methods to the case of problems with partially integer variables, since many tunable algorithms have discrete hyperparameters.
+
+We have tested our newly developed iOpt framework on real datasets, and experimental results show that our method can search for hyperparameters with comparable (and often better) values of the objective metric.
+At the same time, due to its deterministic nature, our method provides a stable result at any run, whereas known randomized algorithms need to be run several times to achieve a good result.
+This ensures a significant superiority of our framework in terms of the time required to select the optimal variant.
+
+In the future, we plan to implement the parallelization of our algorithm, with a focus on heterogeneous computing systems.
+This will allow us to perform hyperparameter tuning of deep neural networks, which requires significant resources and cannot be performed in an acceptable time on a single computing device.
 

### 2023-08-30 2023/MMST/samplepaper.tex
fix some mistakes

@@
 \usepackage{amsmath}
+\usepackage{marvosym}
+\usepackage{hyperref}
 
-\usepackage{hyperref}
+\def\letter{$^{\textrm{(\Letter)}}$}
 
@@
 %
-\title{An algorithm for finding the global extremum 
+\title{An algorithm for finding the global extremum\\ 
 of a partially defined function\thanks{This work was supported by the Russian Science Foundation, project No. 21-11-00204.}}
@@
 %
-\author{Marina Usova\inst{1}\orcidID{0000-0002-0722-6884} \and
+\author{Marina Usova\inst{1}\letter\orcidID{0000-0002-0722-6884} \and
 Konstantin Barkalov\inst{1}\orcidID{0000-0001-5273-2471}}
@@
 \institute{
-Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603022 Nizhni Novgorod, Russia
+Lobachevsky State University of Nizhny Novgorod, \\ Nizhny Novgorod, Russian Federation\\
 \email{\{usova,konstantin.barkalov\}@itmm.unn.ru}}
@@
 
-\keywords{Lipschitz global optimization \and multiextremal functions \and hidden constraints \and dimensionality reduction.}
+\keywords{Lipschitz global optimization \and Multiextremal functions \and Partially defined functions \and Dimensionality reduction}
 \end{abstract}
@@
 \section{Introduction}
+
 Global optimization deals with the development of methods for solving problems of finding the global (absolute) minima of multidimensional multiextremal functions.
-The possibility of constructing a reliable estimate of the global extremum in such problems is principally based on the availability of the \textit{a priori} known information on the problem to be solved.
+The possibility of constructing a reliable estimate of the global extremum in such problems is principally based on the available information on the problem to be solved.
 This information makes it possible to relate the unknown values of the objective function to the known values at the points of the performed search trials.
 
-Often such information is represented in the form of suggestion that the objective function $\phi$ satisfies the Lipschitz condition with a priori unknown constant $L$.
+Often such information is represented in the form of suggestion that the objective function $\phi(y)$ satisfies the Lipschitz condition with a priori unknown constant $L$.
 This suggestion can be interpreted (in the relation to the applied problems) as the reflection of a limited power generating the changes in the simulated system.
-At that, the objective function defined in the form of a ``\textit{black box}'' type functions as a rule is a non differentiable one, and each computation of its value in some point within a feasible domain may require considerable computational resources.
+In this case, the objective function is defined in the form of a ``\textit{black box}'', may be a non-differentiable one, and each computation of its value at some point within a feasible domain may require considerable computational resources.
 
-A number of efficient deterministic algorithms for solving the Lipschitz global optimization problems have been developed ~\cite{Grishagin2018,Jones2021,Birect2020,PaulaviciusZilinskas2014,Sergeyev2017}.
-The comparisons conducted have shown the deterministic algorithms to overcome (in various criteria) the widespread nature-inspired algorithms ~\cite{Liberti2005,Sergeyev2018}.
+A number of efficient deterministic algorithms for solving the Lipschitz global optimization problems have been developed ~\cite{Grishagin2018,Jones2021,PaulaviciusZilinskas2014,Birect2020,Sergeyev2017}.
+The comparisons conducted have shown the deterministic algorithms outperform (in various criteria) the widespread nature-inspired algorithms ~\cite{Liberti2005,Sergeyev2018}.
 
-The present work continues the development of one of the efficient deterministic algorithms for solving the Lipschitz global optimization problems, namely the of the information-statistical global search algorithm ~\cite{Strongin2000,Sergeyev2013}.
+The present work continues the development of one of the efficient deterministic algorithms for solving the Lipschitz global optimization problems, namely the information-statistical global search algorithm ~\cite{Sergeyev2013,Strongin2000}.
 This algorithm suggests the solving of a multidimensional problem by means of its reduction to an equivalent one-dimensional optimization problem using the Peano-Hilbert space-filling curves.
@@
 Note that the use of complex mathematical models is natural in the optimization of complex real world objects that, as a consequence, essentially increases the computation costs for searching the optimum.
-In last decades, the experts in the fields of optimization and parallel computations have proposed many methods for reducing the computation costs an speed-up of the algorithms related to solving the appearing optimization problems \cite{Kvasov2013,Sergeyev2020} as well as to the numerical analysis of the original models \cite{Dongarra2022,Duwe2020}.
+In last decades, the experts in the fields of optimization and parallel computations have proposed many methods for reducing the computation costs and speed-up of the algorithms related to solving the appearing optimization problems \cite{Kvasov2013,Sergeyev2020} as well as to the numerical analysis of the original models \cite{Dongarra2022,Duwe2020}.
 
-However, a principally new problem in applications, namely the numerical instability of the models investigated in some (unknown a priori) regions in the ranges of the parameters’ variations recently became relevant.
-In such regions, one cannot perform the numerical modeling and compute the objective function values correctly.
-This phenomenon can be interpreted either as the presence of some hidden (unknown a priori) constraints \cite{Stripinis2021} or as a partial computability of the objective function in the search domain \cite{Candelieri2019,Sergeyev2003,Strongin2020}.
+However, a principally new problem in applications, namely the numerical instability of the models investigated in some (unknown a priori) subregions of the parameters' variations recently became relevant.
+In such subregions, one cannot perform the numerical modeling and compute the objective function values correctly.
+This phenomenon can be interpreted either as the presence of some hidden constraints \cite{Stripinis2021} or as a partial computability of the objective function in the search domain \cite{Candelieri2019,Sergeyev2003,Strongin2020}.
 In such a statement, the optimization problem becomes more complex since the region of feasible parameters' combinations is not defined in advance.
 
-This paper presents the results in a new research direction related to the use of imputation\footnote{The term ``imputation'' in the paper is used following the established terminology in the field of machine learning related to the restoring of the missing data.} (restoring) of the objective function values in the course of the global search algorithm operation when solving the problems with partially defined objective functions.
+This paper presents the results in a new research direction related to the use of imputation\footnote{The term ``imputation'' in the paper is used following the established terminology in the field of machine learning related to the restoring of the missing data.} (restoring) of the objective function values in the course of the global search algorithm operation when solving the problem with partially defined objective function.
 The approach proposed is based on an adaption of the computation rules of the global search algorithm.
@@
 \begin{equation}\label{eq1} 
-\phi^*=\phi(y^* )=\min_{y \in D} \phi(y), D=\left\{ y \in R^N: a_i \leq y_i \leq b_i, a,b \in R^N,1 \leq i \leq N \right\},
+\phi^*=\phi(y^* )=\min_{y \in D} \phi(y), D=\left\{ y \in R^N: a_i \leq y_i \leq b_i, \; 1 \leq i \leq N \right\},
 \end{equation}
@@
 \begin{equation}\label{eq3} 
-| \phi (y')-\phi (y'') | \leq L \| y'-y'' \|, y',y' \in D,
+| \phi (y')-\phi (y'') | \leq L \| y'-y'' \|, \; y',y'' \in D,
 \end{equation}
@@
 \begin{equation}\label{eq4} 
-| \phi (x')-\phi (x'') | \leq K \| x'-x'' \| ^{1/N}, x',x' \in [0,1],
+| \phi (x')-\phi (x'') | \leq K \| x'-x'' \| ^{1/N}, \; x',x'' \in [0,1],
 \end{equation}
-where $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the rule $K \leq 2L\sqrt {N+3}$ \cite{Strongin2000}.
+where $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the relation $K \leq 2L\sqrt {N+3}$ \cite{Strongin2000}.
 
@@
 The problems of this kind were successfully solved by the proposed algorithm as well but a large number of trials was required to do it since the method attempted to explore the whole non-computability region.
-In this connection, in the rules of constructing the test problems for the dimensionality $N=5$ were modified so that the non-computability region were located olny in the corners of the search domain.}.
+In this connection, in the rules of constructing the test problems for the dimensionality $N=5$ were modified so that the non-computability region were located only in the corners of the search domain.}.
 
-The last assumption makes the application of the classic information-statistical \textit{global search algorithm} \cite{Strongin2000} or other Lipschitz optimization methods \cite{Sergeyev2017} impossible and requires developing a modification for solving the problems with the objective functions, which are not defined everywhere.
+The last assumption makes the application of the information-statistical global search algorithm \cite{Strongin2000} or other Lipschitz optimization methods \cite{PaulaviciusZilinskas2014,Sergeyev2017} impossible and requires developing a modification for solving the problems with the objective functions, which are not defined everywhere.
 
@@
 
-\paragraph{Rule 5.} Select the interval $t$ with the largest value of characteristic $R(i):R(t)= \max_{1 \leq i \leq k+1} R(i)$.
+\paragraph{Rule 5.} Select the interval $(x_{t-1},x_t)$ with the largest value of characteristic $R(i)$:
+\[
+R(t)= \max\{R(i): \; 1 \leq i \leq k+1\}.
+\]
 
@@
 
-\subsection{ Approaches to the Reconstruction of the Objective Function Values }
+\subsection{ Approaches to the Imputation of the Objective Function Values }
 
@@
 
-\subsubsection{The second approach.} To ignore the absence of the objective function values at the point.
+\subsubsection{The second approach.} To ignore the absence of the objective function value at the point.
 The characteristic of the interval with two non-computable point or with one non-computable point and one boundary point is equal to the interval length multiplied by a coefficient, i.e.
@@
 This approach partly resolves the problems of the first approach outlined above.
-However, it doesn’t solve the problem of excess trials and requires a considerable increasing of the parameter $r$ for more complex test problem.
-The examples for the one-dimensional problems using the first two approaches are presented in Fig.~\ref{fig1} and Fig.~\ref{fig2} where the trials falling into the region with the undefined values are marked by the black points and the computable points of the search domain are marked by the green ones.
+However, it doesn't solve the problem of excess trials and requires a considerable increasing of the parameter $r$ for more complex test problem.
+The examples for the one-dimensional problems using the first two approaches are presented in Fig.~\ref{fig1} and Fig.~\ref{fig2}, where the trials falling into the region with the undefined values are marked by the black points and the computable points of the search domain are marked by the green ones.
 
@@
 
-\subsubsection{The fourth approach.} To restore the objective function values in the points for the intervals with two non-computable points as the averages of the nearest computable points taking into account the coordinates, i.e.
+\subsubsection{The fourth approach.} To restore the objective function values at the points of the intervals with two non-computable points as the averages of the nearest computable points taking into account the coordinates, i.e.
 \begin{equation}\label{eq17} 
@@
 \end{equation}
-where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval and $\rho(x_1,x_2) =  |x_1 - x_2|^{1/N}$.
+where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval $(x_{i-1},x_i)$ and $\rho(x_1,x_2) =  |x_1 - x_2|^{1/N}$.
 The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located only near the boundary point.
@@
 \subsubsection{The fifth approach.} To utilize the search information and the suggestions on the function form most completely, one can restore the objective function values in the points taking onto account the H{\"o}lder condition.
-\paragraph{Step 1} Find the end points of the crossing of majorant and minorant
+\paragraph{Step 1.} Find the end points of the crossing of majorant and minorant
 \begin{equation}\label{eq18} 
@@
 \end{equation}
-\paragraph{Step 2} Restore the objective function values according to the rule
+\paragraph{Step 2.} Restore the objective function values according to the rule
 \begin{equation}\label{eq19} 
@@
 
-\begin{figure}[h]
-\includegraphics[width=\textwidth]{fig6.png}
-\caption{Examples of problems No.No. 19, 25, and 35 from the first series of experiments} \label{fig6}
-\end{figure}
-
 \begin{table}[h]
@@
 
-The use of rule (\ref{eq25}) forms a grid of trials inside the non-computability regions that is visible clearly for Problem No.
+The use of rule (\ref{eq25}) forms a grid of trials inside the non-computability regions that is visible clearly for problem No.
 35.
-At that, the trial points are denser at the boundary of the non-computable region close to potential minimum when attempting to approach this one from different sides due to the computational rules (\ref{eq22})-(\ref{eq24}) that was observed in all problems presented as the examples.
-
+In this case, the trial points are denser at the boundary of the non-computable region close to potential minimum when attempting to approach this one from different sides due to the computational rules (\ref{eq22})--(\ref{eq24}) that was observed in all problems presented as the examples.
 100\% of the problems were solved.
@@
 \hline
-$3$ & classical & 12251.1 & - & 100 / 100 \\
+$3$ & original & 12251.1 & - & 100 / 100 \\
 $ $ & round & 12201.8 & 4763.6 & 100 / 100 \\
@@
 \hline
-$4$ & classical & 25144.5 & - & 100 / 100 \\
+$4$ & original & 25144.5 & - & 100 / 100 \\
 $ $ & round & 29840.3 & 15004.4  & 100 / 100 \\
@@
 \hline
-$5$ & classical & 56725.9 & - & 100 / 100 \\
+$5$ & original & 56725.9 & - & 100 / 100 \\
 $ $ & round & 47840.1 & 24109.2 & 100 / 100\\
@@
 In the experiments with the problems of the dimensions $N=3,4,5$, the reliability parameter was set to $r=4.5$, the precision $\varepsilon=0.02$.
-The global minimum was considered to be found if the algorithm generates a trial point inside the $\varepsilon$-nearness of the global minimum.
+The global minimum was considered to be found if the algorithm generates a trial point inside the $\varepsilon$-vicinity of the global minimum.
 The results of the experiments are presented in Table~\ref{tab2}.

### 2023-08-31 2023/MMST/samplepaper.tex
fix last picture (titles) + some titles in text

@@
 \hline
- &  { classical }  & { round district }  & { random district }  \\
+ &  { original }  & { round }  & { random }  \\
 \hline
@@
 \hline
- N &  Task type & Average number of trials & \begin{tabular}{@{}c@{}} {Average number of trials} \\  {in non-computability regions} \end{tabular} & Solved problems \\
+ \begin{tabular}{@{}c@{}} {N} \\ {}\end{tabular} &
+ \begin{tabular}{@{}c@{}} {Problem} \\ {type} \end{tabular} &
+ \begin{tabular}{@{}c@{}} {Average number of trials} \\ {} \end{tabular} &
+ \begin{tabular}{@{}c@{}} {Average number of trials} \\ {in non-computability regions} \end{tabular} &
+ \begin{tabular}{@{}c@{}} {Solved problems} \\ {}\end{tabular} \\
 \hline

### 2023-08-31 2023/MMST/samplepaper.tex
some fixes

@@
 
-First, the objective function may be multiextremal, nondifferentiable, and, moreover, defined in the form of a ``black box'' (i.e.
+First, the objective function may be multiextremal, non-differentiable, and, moreover, defined in the form of a ``black box'' (i.e.
 in the form of some subroutine, which entry the argument is supplied to, and the output is respective function value).
+
 Second, each computing of the function at some point of the feasible region may require considerable computation resources.
@@
   \item Replacing by the most frequent value or by a constant (differs from the use of the median only by the ability to work with the qualitative parameters);
-  \item Replacing the data using the method of $k$ nearest neighbors (unlike the former, accounts for the correlation between the parameters, sensitive to the fadeouts, computation costly);
+  \item Replacing the data using the method of $k$-nearest neighbors (unlike the former, accounts for the correlation between the parameters, sensitive to the fadeouts, computation costly);
   \item Multiple imputation by chained equations (MICE) -- imputation of each value is performed not one but many times (allows understanding how reliable or unreliable is the value suggested);
@@
 
-Each series of experiments consisted of 100 GKLS problems with everywhere computable objective function (the original GKLS problems), 100 problems with the non-computabilities at the boundaries of the search domains, and 100 problems with non-computability regions defined randomly.
+Each series of experiments consisted of 100 GKLS problems with everywhere computable objective function (original), 100 problems with the non-computabilities at the boundaries of the search domains (round), and 100 problems with non-computability regions defined randomly (random).
+
+\begin{figure}[h]
+\includegraphics[width=\textwidth]{fig6.png}
+\caption{Examples of problems No. 19, 25, and 35 from the first series of experiments} \label{fig6}
+\end{figure}
 
@@
 The results of experiments are presented in Table~\ref{tab1}.
+
+Fig.~\ref{fig6} presents the examples of solving several problems with the modified GSA.
+The trials inside the computable regions are marked by the blue points, the global minima of the problems are marked by the red points, and the points of trials performed inside the non-computability regions are marked by the black points.
+Also, the non-computability regions are filled with grey for clarity.
+
+Since the modified algorithm works with the original GKLS problem series according to the rules of base algorithm, one can compare it with the GSA.
+Thus, the general behavior of GSA was preserved for the modification, and the same regions of trials' accumulation near the local minima were observed that can be explained by the realization of the  computation rules (\ref{eq21}) in most cases.
+One can see in Table~\ref{tab1} that the number of trials executed inside the non-computable regions is small.
+Total number of trials for the original problems and for the problems with partially defined objective function are comparable, the method doesn't perform excess trials in the non-promising regions due to the computing formulae utilizing the information on the nearest computable points.
 
@@
 \end{table}
-
-\begin{figure}[h]
-\includegraphics[width=\textwidth]{fig6.png}
-\caption{Examples of problems No. 19, 25, and 35 from the first series of experiments} \label{fig6}
-\end{figure}
-
-Fig.~\ref{fig6} presents the examples of solving several problems with the modified GSA.
-The trials inside the computable regions are marked by the blue points, the global minima of the problems are marked by the red points, and the points of trials performed inside the non-computability regions are marked by the black points.
-Also, the non-computability regions are filled with grey for clarity.
-
-Since the modified algorithm works with the original GKLS problem series according to the rules of base algorithm, one can compare it with the GSA.
-Thus, the general behavior of GSA was preserved for the modification, and the same regions of trials' accumulation near the local minima were observed that can be explained by the realization of the  computation rules (\ref{eq21}) in most cases.
-One can see in Table~\ref{tab1} that the number of trials executed inside the non-computable regions is small.
-Total number of trials for the original problems and for the problems with partially defined objective function are comparable, the method doesn't perform excess trials in the non-promising regions due to the computing formulae utilizing the information on the nearest computable points.
 

### 2023-08-31 2023/vestnik.susu/example.tex
fix abstract

@@
 
-\title{ON USING THE DECISION TREES TO IDENTIFY THE LOCAL EXTREMA IN PARALLEL GLOBAL OPTIMIZATION ALGORITHMS
+\title{ON USING THE DECISION TREES TO IDENTIFY THE LOCAL EXTREMA IN PARALLEL GLOBAL OPTIMIZATION ALGORITHM
 \footnote{The paper is recommended for publication by the Program Committee of the International Scientific Conference “Parallel Computational Technologies (PCT) 2023”}
@@
 	In the present work, the solving of the multidimensional global optimization problems using decision tree to reveal the attractor regions of the local minima is considered.
-The objective function of the  problem is defined as a ``black box''.
-We assume it only to satisfy Lipschitz condition with unknown constant.
+The objective function of the  problem is defined as a ``black box'', may be non-differentiable, multi-extremal and computational costly.
+We assume that the function satisfies the Lipschitz condition with a priory unknown constant.
 Global search algorithm is applied for the search of global minimum in the problems of such type.
-In the present work, we propose a method for selecting the vicinity of local extrema of the objective  function based on analysis of accumulated search information.
+It is well known that the solution complexity essentially depends on the presence of multiple local extrema.
+Within the framework of the global search algorithm, we propose a method for selecting the vicinity of local extrema of the objective  function based on analysis of accumulated search information.
 Conducting such an analysis using machine learning techniques allows making a decision to run a local  method, which can speed up the convergence of the algorithm.
@@
 
-\forcitationen{Pervyi~A.B., Vtoroi~V.G. The \LaTeXe{} Template for Paper Submission for the ``Computational Mathematics and Software Engineering'' Series}
+\forcitationen{Barkalov~K.A., Lebedev~I.G., Silenko D.I. On using the decision trees to identify the local extrema in parallel global optimization algorithm}
 
@@
 In Section \ref{SecA}, the description of the methods and approaches used is given.
-In particular, in Section \ref{SecGSA},  the main idea of global search algorithm is discussed, its computational rules are given.
-Section \ref{SecHG} contains the description of the local search method used.
-The method of constructing the approximation of the function using the decision trees is described in  Section \ref{SecDT}.
+In particular, the main idea of global search algorithm is discussed and its computational rules are given.
+The section also contains the description of the local search method used and the method of constructing the approximation of the function using the decision trees.
 The novel algorithm combining the global search and the local one on the base of using the decision  trees is presented in Section \ref{SecGSAL}.

### 2023-09-01 2023/ITMO paper/elsarticle-template-num.tex
translate one sectence in Rissuan

@@
 %% Русский язык - закомментировать в финальной версии
-\usepackage[russian]{babel}
+%\usepackage[russian]{babel}
 
@@
 
-The simplest way to find the global optimum is the exhaustive search method (поиска по равномерной \cite{Bao2006} или случайной \cite{Bergstra2012} сетке).
+The simplest way to find the global optimum is the exhaustive search method (search on the uniform \cite{Bao2006} or random \cite{Bergstra2012} grid). 
 By selecting the number of analyzed variants for each of the hyperparameters, searching all the resulting combinations and choosing the best of them in a given metric, we obtain an estimate of the optimum with a certain accuracy \cite{Nevendra2022}.

### 2023-09-01 2023/ITMO paper/elsarticle-template-num.tex
minor fix

@@
 
-The simplest way to find the global optimum is the exhaustive search method (search on the uniform \cite{Bao2006} or random \cite{Bergstra2012} grid). 
+The simplest way to find the global optimum is the exhaustive search method (search on a uniform \cite{Bao2006} or random \cite{Bergstra2012} grid). 
 By selecting the number of analyzed variants for each of the hyperparameters, searching all the resulting combinations and choosing the best of them in a given metric, we obtain an estimate of the optimum with a certain accuracy \cite{Nevendra2022}.
 The exhaustive search method is easy to implement, it has perfect parallelism, but it has a significant drawback (an exponential growth of the number of grid nodes as the number of hyperparameters increases).
-Thus, exhaustive search can be used as a method for solving HPO problems only in "simple" problems where both the number of hyperparameters (one or two) and the time for solving the initial problem with each selected combination of hyperparameters are small.
+Thus, exhaustive search can be used as a method for solving HPO problems only in ``simple'' problems where both the number of hyperparameters (one or two) and the time for solving the initial problem with each selected combination of hyperparameters are small.
 
@@
 Known deterministic methods for solving problems of this class are centered on solving linear or convex problems \cite{Lee2012}.
-It is suggested that complex multi-extremal problems should be solved using heuristic approaches \cite{Belotti2013}.
+It is suggested that complex multiextremal problems should be solved using heuristic approaches \cite{Belotti2013}.
 
@@
 \section*{Code availability}
-%проверить перевод
-Software that implements all of the described methods is available in the open repository \url{https://github.com/aimclub/iOpt/}.
+The software that implements all of the described methods is available in the open repository \url{https://github.com/aimclub/iOpt/}.
 
 \section*{Acknowledgements}
-%уточнить ссылку
-This research was supported by the research center ``Strong Artificial Intelligence in Industry'' of ITMO University.
+This work was supported by the Analytical Center for the Government of the Russian Federation (IGK 000000D730321P5Q0002), agreement No. 70-2021-00141.
 

### 2023-09-02 2023/ITMO paper/elsarticle-template-num.tex
fix some bugs

@@
 %% keywords here, in the form: keyword \sep keyword
-Lipschitz optimization \sep Multiextremal function \sep Hyperparameter tuning
+Lipschitz optimization \sep Global optimization \sep Multiextremal functions \sep Hyperparameter tuning
 
@@
 
-In this paper, we present iOpt \footnote{\url{https://github.com/aimclub/iOpt/}}, a framework of intelligent optimization methods focused on solving problems of selecting optimal values of hyperparameters of various algorithms (including machine learning and heuristic optimization methods).
+In this paper, we present iOpt\footnote{\url{https://github.com/aimclub/iOpt/}}, a framework of intelligent optimization methods focused on solving problems of selecting optimal values of hyperparameters of various algorithms (including machine learning and heuristic optimization methods).
 The paper is organized as follows.
@@
 \end{minipage}
-\caption{Evolvents $y_m(x)$ with $m=4$.}\label{fig:Peano}
+\caption{Evolvents $y_m(x)$ with $m=4$ for dimensions $N=2$ and $N=3$.}\label{fig:Peano}
 \end{figure}
 
-Remark 2: The theoretical Peano-Hilbert curve $y(x)$ is defined as a limiting object.
+Remark 2.
+The theoretical Peano-Hilbert curve $y(x)$ is defined as a limiting object.
 Numerical algorithms use evolvents $y_m(x)$, that approximate the true curve with a given density level $2^{-m}$, which depends on the required search accuracy.
@@
 The iOpt framework was designed in the Python programming language using its standard library.
-To use the framework, you need to have the Pyhon interpreter version 3.8 or higher installed.
+To use the framework, you need to have the Python interpreter version 3.8 or higher installed.
 The framework is open source and is available at \url{https://github.com/aimclub/iOpt}.
@@
  \KwOut{$y_{min}$}
- Initialize $k=2$, $\Omega_k= \left\{ (x_1=0, y_1=y(x_1), z_1=\varphi(y_1)), (x_2=1, y_2=y(x_2), z_2=\varphi(y_2)) \right\}$\\
+ Initialize $k=2$, $t=2$, $\Omega_k= \left\{ (x_1=0, y_1=y(x_1), z_1=\varphi(y_1)), (x_2=1, y_2=y(x_2), z_2=\varphi(y_2)) \right\}$\\
  \While{ $\Delta_t \geq \epsilon$ and  $k \leq K_{max}$}{
@@
 
-\subsection{Experimental setup}
 In this section, we present experimental results that demonstrate the effectiveness of the developed framework for a number of HPO problems.
@@
 
+\subsection{Experimental setup}
+
 \textbf{Datasets.} 
-We used for our experiments fifteen datasets from the UCI repository \cite{AsuncionNewman2007}, which covers a large number of areas.
-Datasets from this repository are quite often used to test the correct operation of hyperparameter tuning algorithms \cite{WU2023110160,Wang2021,Joy2020}.
+We used for our experiments fifteen datasets from the UCI repository \cite{UCI}, which covers a large number of areas.
+Datasets from this repository are quite often used to test the correct operation of hyperparameter tuning algorithms \cite{Joy2020,Wang2021,Wu2023}.
 We selected the datasets to ensure that the data contained in them were described by both numerical and categorical attributes.
@@
 \label{tab:1}
-\begin{tabular}{lllllll}
+\begin{tabular}{lcccccc}
 \hline
@@
 The implementation of the method is based on the widely used XGBoost library.
- \item Multi-Layer Perceptron (MLP), a fully connected artificial neural network, is a supervised learning algorithm that forms a functionf:$R^m \to R^o$, where $m$ is the number of input attributes, and o is the output dimensionality of the network \cite{GARDNER19982627}.
+ \item Multi-Layer Perceptron (MLP), a fully connected artificial neural network, is a supervised learning algorithm that forms a function $f : R^m \to R^o$, where $m$ is the number of input attributes, and $o$ is the output dimensionality of the network \cite{GARDNER19982627}.
 Using this method, it is possible to form the function f as a universal nonlinear approximator for the classification and regression problem.
 In the framework of this study, we will consider the multilayer perceptron for the task of attributing an object to a particular class.
- \item Support Vector Classification (SVC) is a support vector method for the classification problem \cite{SVC2003}.
-The main purpose of this algorithm is to find the equation of a separating hyperplane $(\sum_{i=1}^n \omega_i x_i)+\omega_0=0$ in the space $R^n$ that separates classes in some optimal way.
+ \item Support Vector Classification (SVC) is a support vector method for the classification problem \cite{SVC}.
+The main purpose of this algorithm is to find the equation of a separating hyperplane $\omega_0 + \sum_{i=1}^n \omega_i x_i=0$ in the space $R^n$ that separates classes in some optimal way.
 To improve the predictive ability of the algorithm, a method for constructing a nonlinear classifier was proposed by switching from scalar products to arbitrary kernels, which allows the construction of nonlinear separators.
-An implementation of the SVC method from the scikit learn library was used to investigate the capabilities of HPO algorithms.
+An implementation of the SVC method from the scikit-learn library was used to investigate the capabilities of HPO algorithms.
 
@@
 \label{tab:2}
-\begin{tabular}{llll}
+\begin{tabular}{lccc}
 \hline
@@
 This metric has been applied when working with the datasets Optical Recognition of Handwritten Digits, Bank Marketing, CNAE9, Statlog Segmentation, Semeion Handwritten Digit;
- \item Binary F1 - F1 score is adapted for binary classification tasks, with the predominance of one class over another in the dataset \cite{Chicco2020}.
+ \item Binary F1 -- F1 score is adapted for binary classification tasks, with the predominance of one class over another in the dataset \cite{Chicco2020}.
 This metric is suitable for analyzing Breast Cancer Wisconsin, Credit Approval, Parkinsons, Banknote authentication, Wilt datasets;
- \item Macro F1 - F1 score is suitable for tasks with a large number of unbalanced classes \cite{Metric2}.
+ \item Macro F1 -- F1 score is suitable for tasks with a large number of unbalanced classes \cite{Metric2}.
 The datasets Protein Localization Sites (Ecoli), Balance Scale Weight \& Distance, Zoo, Car Evaluation, Turkiye Student Evaluation have this particular feature.
@@
 There are currently three algorithms implemented in Hyperopt: Random Search, Tree of Parzen Estimators (TPE) and Adaptive TPE.
-We used the computation variant with the TPE algorithm implementation - Hyperopt(tpe) \cite{NIPS2011_86e8f7ab}.
+We used the computation variant with the TPE algorithm implementation -- Hyperopt(tpe) \cite{NIPS2011_86e8f7ab}.
 

### 2023-09-06 2023/ITMO paper/elsarticle-template-num.tex
minor corrections

@@
 
-\author[UNN,ITMO]{Konstantin Barkalov}
+\author[UNN,ITMO]{Konstantin Barkalov\corref{cor}}
+\cortext[cor]{Corresponding author: \url{konstantin.barkalov@itmm.unn.ru}}
 \author[UNN,ITMO]{Denis Karchkov}
@@
 
-The iOpt framework of intelligent optimization methods was developed as part of the ``Strong Artificial Intelligence in Industry'' project.
-It is designed to select optimal (in a given metric) values of parameters of complex objects and processes, for example, artificial intelligence and machine learning methods.
+The iOpt framework of intelligent optimization methods is designed to select optimal (in a given metric) values of parameters of complex objects and processes, especially artificial intelligence and machine learning methods.
 The framework allows fine-tuning of parameters of models and methods used in applied research in various fields of science.
@@
 The software that implements all of the described methods is available in the open repository \url{https://github.com/aimclub/iOpt/}.
+Python scripts that reproduce the results of the presented experiments are available at \url{https://github.com/aimclub/iOpt/benchmarks}.
 

### 2023-09-06 2023/ITMO paper/elsarticle-template-num.tex
fix two prepositions

@@
 
-The iOpt framework of intelligent optimization methods is designed to select optimal (in a given metric) values of parameters of complex objects and processes, especially artificial intelligence and machine learning methods.
+The iOpt framework of intelligent optimization methods is designed to select optimal (in a given metric) values of parameters of complex objects and processes, especially in artificial intelligence and machine learning methods.
 The framework allows fine-tuning of parameters of models and methods used in applied research in various fields of science.
@@
 The software that implements all of the described methods is available in the open repository \url{https://github.com/aimclub/iOpt/}.
-Python scripts that reproduce the results of the presented experiments are available at \url{https://github.com/aimclub/iOpt/benchmarks}.
+Python scripts that reproduce the results of the experiments described in this paper are available at \url{https://github.com/aimclub/iOpt/benchmarks}.
 

### 2023-09-29 2023/vestnik.susu/example.tex
Допереведена аннотация

@@
 	В работе рассматривается решение многомерных задач многоэкстремальной оптимизации с использованием деревьев решений для выявления областей притяжения локальных минимумов.
-Целевая функция представлена как «черный ящик»; для нее лишь предполагается, что она удовлетворяет условию Липшица с априори неизвестной константой.
-Для решения задачи многоэкстремальной оптимизации применятся алгоритм глобального поиска.
-Предложена модификация алгоритма, в которой определяются окрестности локальных минимумов целевой функции на основе анализа накопленной поисковой информации.
+Целевая функция представлена как «черный ящик», она может быть недифференцируемой, многоэкстремальной и вычислительно трудоемкой.
+Для функции предполагается, что она удовлетворяет условию Липшица с априори неизвестной константой.
+Для решения поставленной задачи многоэкстремальной оптимизации применятся алгоритм глобального поиска.
+Хорошо известно, что сложность решения существенно зависит от наличия нескольких локальных экстремумов.
+В данной работе предложена модификация алгоритма, в которой определяются окрестности локальных минимумов целевой функции на основе анализа накопленной поисковой информации.
 Проведение такого анализа с использованием методов машинного обучения позволяет принять решение о запуске локального метода, что может ускорить сходимость алгоритма.

### 2023-09-29 2023/vestnik.susu/example.tex
Исправлены замечания корректора.

@@
 }
+
+\selectlanguage{english}
 
@@
 Since in real global optimization problems each computing of the function value (hereafter called  \textit{a trial}) is a computation-costly operation, one has to reduce the number of such trials.
-It  can be achieved by intentional choice of variants in the course of search for the optimal solution cutting  off unpromising search subdomains and investigating only these ones, in which the solution of the  problem can be found.
+It  can be achieved by intentional choice of variants in the course of search for the optimal solution cutting  off unpromising search subdomains and investigating only the ones, in which the solution of the  problem can be found.
 The global search algorithm (GSA) is based on this idea \bblcite{Strongin2000}.
@@
 
-In order to determine,  at which moment it is best to run the local method within the framework of the  global search, we will use an algorithm based on a decision tree.
+In order to determine  at which moment it is best to run the local method within the framework of the  global search, we will use an algorithm based on a decision tree.
 At the first search stage, the search information is accumulated in the form of the results of the trials  executed.
@@
 
-Also, assume the function to satisfy the Lipschitz condition with \textit{a priori} unknown constant $L$, that corresponds to limited variation of the function values at limited variation of the argument.
+Also, assume the function to satisfy the Lipschitz condition with \textit{a priori} unknown constant $L$ that corresponds to limited variation of the function values at limited variation of the argument.
 \begin{displaymath} 
@@
 
-The Hooke--Jeeves pattern search algorithm belongs to the class of zero-order methods.
+For the Hooke--Jeeves pattern search algorithm belongs to the class of zero-order methods.
 Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \bblcite{Himmelblau72}.
@@
 			\includegraphics[width=1\linewidth]{figure/fig1.pdf} 
-			\caption{An example of iterations of the Hooke-Jeeves algorithm} %%  подпись к рисунку 
+			\caption{Example of iterations of the Hooke-Jeeves algorithm} %%  подпись к рисунку 
 			\label{fig:fig1} 
@@
 	\item	The step value is determined (it is different for each coordinate and can be varied in the course  of search).  
-	\item	The search step is considered to be successful if the value of the objective function at the check  point doesn't exceed the value of the objective function at the initial point.  
+	\item	The search step is considered to be successful if the value of the objective function at the check  point does not exceed the value of the objective function at the initial point.  
 	\item	Otherwise, it is necessary to return to the previous item and make a step in the reverse direction.  
@@
 
-Fig.
-\ref{fig:fig1} shows the example of the algorithm work.
+Figure \ref{fig:fig1} shows the example of the algorithm work.
 Function level lines are displayed, filled circles indicate successful steps, unfilled circles correspond to unsuccessful steps performed during the investigating search.
@@
 Let us denote the value  computed by the  decision tree at the point $y$ as $z' = \psi(y)$.
-Fig. \ref{fig:fig2} presents a graph of the objective function $\varphi(y)$; Fig. \ref{fig:fig2_2} presents corresponding approximation $\psi(y)$ built using decision tree.
+Figure \ref{fig:fig2} presents a graph of the objective function $\varphi(y)$; Figure \ref{fig:fig2_2} presents corresponding approximation $\psi(y)$ built using decision tree.
 
@@
 			\includegraphics[width=1\linewidth]{figure/fig5.png}
-			\caption{An objective function $\varphi(y)$ from the test class} %% подпись к рисунку
+			\caption{Objective function $\varphi(y)$ from the test class} %% подпись к рисунку
 			\label{fig:fig2}
@@
 			\includegraphics[width=1\linewidth]{figure/fig4.png}
-			\caption{A piecewise-constant  approximation $\psi(y)$ of the function $\varphi(y)$} %% подпись к рисунку
+			\caption{Piecewise-constant  approximation $\psi(y)$ of the function $\varphi(y)$} %% подпись к рисунку
 			\label{fig:fig2_2}
@@
 
-In the current section, let us present a detailed description how we used the decision tree for finding  the attraction areas of the local extrema.
+In the current section, let us present a detailed description of how we used the decision tree for finding  the attraction areas of the local extrema.
 In the course of running GSA, it is necessary to determine whether it is worth to use current point as a  start one for the local method or not.
@@
 After completing a certain number of trials (for example, $100\cdot N$), let us use all accumulated  points to initialize a suitable data structure and training the decision tree on its base.
-To make the  determining the adjacent points easier, we construct a uniform grid with certain step in each coordinate.
+To make the  determining of the adjacent points easier, we construct a uniform grid with certain step in each coordinate.
 Next, we compute the values of the approximation of the objective function in these points  using the decision tree.
@@
 This  closest point is a projection of the initial point onto the uniform grid, on which its neighbors can be  determined easily.
-Since the decision tree constructs the approximation of the initial function (even  piecewise constant), we can estimate the function values in the areas, in which the trials haven't been  executed earlier.
+Since the decision tree constructs the approximation of the initial function (even  piecewise constant), we can estimate the function values in the areas, in which the trials have not been  executed earlier.
 
@@
 If even a single  neighbor has a lower function value than current point, there is no need to run the local method.
-If at  even a single neighboring point the function value is the same as at the current one, this point must be also checked (i.e.
-to find all its neighbors and to check these ones in the same way).
+If at  even a single neighboring point the function value is the same as at the current one, this point must be also checked (i.e., to find all its neighbors and to check these ones in the same way).
 Only in the case when  all function values at the neighboring points are greater than at the current point, we can run the local  method.
@@
 Next, the master process selects a new point  $x^{k+p+1}$ for the slave process according to the rules provided below.
-Note that in this case it will be a set of inverse images of the trial points  $I_k=\left\{x^{k+1},x^{k+2},\ldots,x^{k+p}\right\}$, in which the trials have been already initiated but haven't been completed yet.
+Note that in this case it will be a set of inverse images of the trial points  $I_k=\left\{x^{k+1},x^{k+2},\ldots,x^{k+p}\right\}$, in which the trials have been already initiated but have not been completed yet.
 
@@
 	where $ z_i=\varphi\left(y\left(x_i\right)\right)$ if $x_i\notin I_k,\;1\le i\le k+p$.
-The values $z_i$ at the points $x_i\in I_k$ are not defined as far as the trials at these points haven't been completed yet.
+The values $z_i$ at the points $x_i\in I_k$ are not defined as far as the trials at these points have not been completed yet.
 If $M=0$, then set $M=1$.
@@
 
-	If no one neighbor point has a lower value than in $y'_q$, run the local method.
+	If not one neighbor point has a lower value than in $y'_q$, run the local method.
 
@@
 % переводить до этого раздела
-\section{Numerical Experiments}\label{SecR}
+\section{Numerical experiments}\label{SecR}
 

### 2023-09-30 2023/ITMO paper/elsarticle-template-num.tex
add table and fig. for time

@@
 \usepackage{multirow}
+
+\usepackage{graphicx}
+\usepackage[table,xcdraw]{xcolor}
+\usepackage{amssymb}
 
@@
 
+% Please add the following required packages to your document preamble:
+% \usepackage{graphicx}
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+\begin{table}[]
+\centering
+\caption{Values of optimized metrics for the algorithm's averaged over 10 runs}
+\label{tab:res}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{ccccccc}
+\hline
+\textbf{Dataset}                & \textbf{iOpt}                         & \textbf{Optuna(random)}                        & \textbf{Optuna(tpe)}                           & \textbf{Optuna(cmaes)}                         & \textbf{Optuna(nsgaii)}                        & \textbf{Hyperopt(tpe)}                         \\ \hline
+                                & \multicolumn{6}{c}{\textbf{SVM}}                                                                                                                                                                                                                                                           \\ \hline
+\textbf{Balance}                & {\color[HTML]{FE0000} 0.856}          & 0.856 $\pm$ 0.07\%                                 & 0.856 $\pm$ 0.23\%                                 & 0.834 $\pm$ 8.81\%                                 & \textbf{0.857 $\pm$ 0.31\%}                        & 0.856 $\pm$ 0.07\%                                 \\
+\textbf{Bank-marketing}         & {\color[HTML]{FE0000} 0.894}          & 0.880 $\pm$ 1.65\%                                 & 0.892 $\pm$ 2.11\%                                 & 0.861 $\pm$ 5.16\%                                 & 0.879 $\pm$ 3.45\%                                 & \textbf{0.896 $\pm$ 0.48\%}                        \\
+\textbf{Banknote}               & {\color[HTML]{FE0000} \textbf{0.992}} & 0.990 $\pm$ 0.04\%                                 & 0.991 $\pm$ 0.07\%                                 & 0.990 $\pm$ 0.20\%                                 & 0.990 $\pm$ 0.02\%                                 & 0.990 $\pm$ 0.00\%                                 \\
+\textbf{Breast Cancer}          & {\color[HTML]{FE0000} 0.971}          & 0.970 $\pm$ 0.13\%                                 & 0.972 $\pm$ 0.19\%                                 & \textbf{0.973 $\pm$ 0.25\%}                        & 0.972 $\pm$ 0.13\%                                 & \textbf{0.973 $\pm$ 0.26\%}                        \\
+\textbf{CarEvaluation}          & 0.349                                 & 0.350 $\pm$ 4.83\%                                 & {\color[HTML]{FE0000} \textbf{0.376 $\pm$ 4.57\%}} & 0.349 $\pm$ 4.83\%                                 & 0.355 $\pm$ 3.77\%                                 & 0.366 $\pm$ 3.21\%                                 \\
+\textbf{CNAE-9}                 & {\color[HTML]{FE0000} \textbf{0.941}} & 0.938 $\pm$ 0.07\%                                 & 0.939 $\pm$ 0.05\%                                 & 0.939 $\pm$ 0.00\%                                 & 0.938 $\pm$ 0.05\%                                 & 0.939 $\pm$ 0.07\%                                 \\
+\textbf{Credit   Approval}      & 0.878                                 & 0.881 $\pm$ 0.22\%                                 & {\color[HTML]{FE0000} \textbf{0.884 $\pm$ 0.18\%}} & 0.880 $\pm$ 0.23\%                                 & 0.880 $\pm$ 0.22\%                                 & 0.881 $\pm$ 0.21\%                                 \\
+\textbf{Digits}                 & 0.984                                 & 0.984 $\pm$ 0.06\%                                 & {\color[HTML]{FE0000} \textbf{0.985 $\pm$ 0.02\%}} & \textbf{0.985 $\pm$ 0.05\%}                        & 0.984 $\pm$ 0.06\%                                 & \textbf{0.985 $\pm$ 0.03\%}                        \\
+\textbf{Ecoli}                  & {\color[HTML]{FE0000} \textbf{0.863}} & 0.855 $\pm$ 0.28\%                                 & 0.861 $\pm$ 0.45\%                                 & 0.860 $\pm$ 0.59\%                                 & 0.855 $\pm$ 0.15\%                                 & 0.859 $\pm$ 0.28\%                                 \\
+\textbf{Parkinsons}             & {\color[HTML]{FE0000} \textbf{0.931}} & 0.929 $\pm$ 0.33\%                                 & 0.930 $\pm$ 0.28\%                                 & 0.930 $\pm$ 0.19\%                                 & \textbf{0.931 $\pm$ 0.39\%}                        & \textbf{0.931 $\pm$ 0.36\%}                        \\
+\textbf{Semeion}                & {\color[HTML]{FE0000} \textbf{0.942}} & 0.941 $\pm$ 0.10\%                                 & 0.941 $\pm$ 0.19\%                                 & \textbf{0.942 $\pm$ 0.12\%}                        & 0.941 $\pm$ 0.10\%                                 & \textbf{0.942 $\pm$ 0.09\%}                        \\
+\textbf{Statlog   Segmentation} & {\color[HTML]{FE0000} \textbf{0.958}} & 0.955 $\pm$ 0.13\%                                 & 0.957 $\pm$ 0.07\%                                 & 0.956 $\pm$ 0.08\%                                 & 0.956 $\pm$ 0.06\%                                 & 0.956 $\pm$ 0.06\%                                 \\
+\textbf{Wilt}                   & {\color[HTML]{FE0000} \textbf{0.704}} & 0.681 $\pm$ 1.90\%                                 & 0.699 $\pm$ 1.95\%                                 & 0.659 $\pm$ 17.91\%                                & 0.674 $\pm$ 1.89\%                                 & 0.696 $\pm$ 1.07\%                                 \\
+\textbf{Zoo}                    & {\color[HTML]{FE0000} \textbf{0.979}} & {\color[HTML]{FE0000} \textbf{0.979 $\pm$ 0.00\%}} & {\color[HTML]{FE0000} \textbf{0.979 $\pm$ 0.00\%}} & {\color[HTML]{FE0000} \textbf{0.979 $\pm$ 0.00\%}} & {\color[HTML]{FE0000} \textbf{0.979 $\pm$ 0.00\%}} & {\color[HTML]{FE0000} \textbf{0.979 $\pm$ 0.00\%}} \\ \hline
+\textbf{}                       & \multicolumn{6}{c}{\textbf{XGBoost}}                                                                                                                                                                                                                                                       \\ \hline
+\textbf{Balance}                & 0.633                                 & 0.638 $\pm$ 1.16\%                                 & 0.650 $\pm$ 3.26\%                                 & 0.650 $\pm$ 1.73\%                                 & 0.645 $\pm$ 1.50\%                                 & {\color[HTML]{FE0000} \textbf{0.661 $\pm$ 2.63\%}} \\
+\textbf{Bank-marketing}         & 0.918                                 & 0.918 $\pm$ 0.04\%                                 & {\color[HTML]{FE0000} \textbf{0.919 $\pm$ 0.04\%}} & 0.918 $\pm$ 0.04\%                                 & 0.918 $\pm$ 0.03\%                                 & 0.918 $\pm$ 0.03\%                                 \\
+\textbf{Banknote}               & 0.994                                 & 0.994 $\pm$ 0.07\%                                 & \textbf{0.996 $\pm$ 0.07\%}                        & {\color[HTML]{FE0000} \textbf{0.996 $\pm$ 0.08\%}} & 0.995 $\pm$ 0.08\%                                 & 0.995 $\pm$ 0.09\%                                 \\
+\textbf{Breast Cancer}          & 0.961                                 & 0.957 $\pm$ 0.27\%                                 & {\color[HTML]{FE0000} \textbf{0.963 $\pm$ 0.16\%}} & 0.961 $\pm$ 0.17\%                                 & 0.960 $\pm$ 0.39\%                                 & 0.960 $\pm$ 0.25\%                                 \\
+\textbf{CarEvaluation}          & {\color[HTML]{FE0000} 0.980}          & 0.965 $\pm$ 0.88\%                                 & \textbf{0.983 $\pm$ 0.39\%}                        & 0.980 $\pm$ 0.28\%                                 & 0.974 $\pm$ 0.48\%                                 & 0.977 $\pm$ 0.38\%                                 \\
+\textbf{CNAE-9}                 & 0.932                                 & 0.924 $\pm$ 0.83\%                                 & {\color[HTML]{FE0000} \textbf{0.937 $\pm$ 0.34\%}} & 0.935 $\pm$ 0.26\%                                 & 0.930 $\pm$ 0.53\%                                 & 0.935 $\pm$ 0.29\%                                 \\
+\textbf{Credit   Approval}      & {\color[HTML]{FE0000} \textbf{0.901}} & 0.897 $\pm$ 0.29\%                                 & \textbf{0.901 $\pm$ 0.30\%}                        & 0.898 $\pm$ 0.22\%                                 & 0.899 $\pm$ 0.32\%                                 & 0.898 $\pm$ 0.28\%                                 \\
+\textbf{Digits}                 & 0.975                                 & 0.973 $\pm$ 0.17\%                                 & 0.976 $\pm$ 0.15\%                                 & {\color[HTML]{FE0000} \textbf{0.978 $\pm$ 0.15\%}} & 0.975 $\pm$ 0.15\%                                 & 0.976 $\pm$ 0.15\%                                 \\
+\textbf{Ecoli}                  & {\color[HTML]{FE0000} \textbf{0.865}} & 0.852 $\pm$ 0.55\%                                 & \textbf{0.865 $\pm$ 0.53\%}                        & 0.864 $\pm$ 0.24\%                                 & 0.856 $\pm$ 0.40\%                                 & 0.861 $\pm$ 0.37\%                                 \\
+\textbf{Parkinsons}             & {\color[HTML]{FE0000} 0.950}          & 0.944 $\pm$ 0.24\%                                 & \textbf{0.953 $\pm$ 0.42\%}                        & 0.952 $\pm$ 0.16\%                                 & 0.949 $\pm$ 0.32\%                                 & \textbf{0.953 $\pm$ 0.42\%}                        \\
+\textbf{Semeion}                & 0.940                                 & 0.939 $\pm$ 0.48\%                                 & {\color[HTML]{FE0000} \textbf{0.951 $\pm$ 0.20\%}} & 0.950 $\pm$ 0.14\%                                 & 0.942 $\pm$ 0.39\%                                 & 0.948 $\pm$ 0.25\%                                 \\
+\textbf{Statlog   Segmentation} & 0.982                                 & 0.981 $\pm$ 0.11\%                                 & \textbf{0.984 $\pm$ 0.24\%}                        & 0.984 $\pm$ 0.07\%                                 & 0.983 $\pm$ 0.13\%                                 & {\color[HTML]{FE0000} \textbf{0.984 $\pm$ 0.08\%}} \\
+\textbf{Wilt}                   & 0.868                                 & 0.842 $\pm$ 1.41\%                                 & {\color[HTML]{FE0000} \textbf{0.872 $\pm$ 0.32\%}} & 0.867 $\pm$ 0.43\%                                 & 0.858 $\pm$ 0.88\%                                 & 0.865 $\pm$ 0.34\%                                 \\
+\textbf{Zoo}                    & {\color[HTML]{FE0000} \textbf{1.000}} & 0.992 $\pm$ 0.86\%                                 & 0.992 $\pm$ 0.79\%                                 & 0.989 $\pm$ 0.73\%                                 & 0.998 $\pm$ 0.47\%                                 & 0.987 $\pm$ 0.64\%                                 \\ \hline
+\textbf{}                       & \multicolumn{6}{c}{\textbf{MLP}}                                                                                                                                                                                                                                                           \\ \hline
+\textbf{Balance}                & 0.968                                 & 0.967 $\pm$ 0.41\%                                 & {\color[HTML]{FE0000} \textbf{0.973 $\pm$ 0.30\%}} & 0.966 $\pm$ 0.39\%                                 & 0.969 $\pm$ 0.32\%                                 & 0.970 $\pm$ 0.32\%                                 \\
+\textbf{Bank-marketing}         & 0.914                                 & 0.914 $\pm$ 0.05\%                                 & \textbf{0.915 $\pm$ 0.04\%}                        & 0.914 $\pm$ 0.03\%                                 & {\color[HTML]{FE0000} \textbf{0.915 $\pm$ 0.03\%}} & {\color[HTML]{FE0000} \textbf{0.915 $\pm$ 0.03\%}} \\
+\textbf{Banknote}               & {\color[HTML]{FE0000} \textbf{1.000}} & {\color[HTML]{FE0000} \textbf{1.000 $\pm$ 0.00\%}} & {\color[HTML]{FE0000} \textbf{1.000 $\pm$ 0.00\%}} & {\color[HTML]{FE0000} \textbf{1.000 $\pm$ 0.00\%}} & {\color[HTML]{FE0000} \textbf{1.000 $\pm$ 0.00\%}} & {\color[HTML]{FE0000} \textbf{1.000 $\pm$ 0.00\%}} \\
+\textbf{Breast Cancer}          & {\color[HTML]{FE0000} \textbf{0.974}} & \textbf{0.974 $\pm$ 0.16\%}                        & \textbf{0.974 $\pm$ 0.07\%}                        & 0.973 $\pm$ 0.12\%                                 & \textbf{0.974 $\pm$ 0.09\%}                        & \textbf{0.974 $\pm$ 0.11\%}                        \\
+\textbf{CarEvaluation}          & 0.996                                 & 0.999 $\pm$ 0.10\%                                 & 0.993 $\pm$ 1.33\%                                 & 0.982 $\pm$ 1.98\%                                 & \textbf{1.000 $\pm$ 0.03\%}                        & {\color[HTML]{FE0000} \textbf{1.000 $\pm$ 0.00\%}} \\
+\textbf{CNAE-9}                 & 0.957                                 & 0.955 $\pm$ 0.12\%                                 & 0.958 $\pm$ 0.06\%                                 & 0.955 $\pm$ 0.14\%                                 & 0.957 $\pm$ 0.15\%                                 & {\color[HTML]{FE0000} \textbf{0.959 $\pm$ 0.17\%}} \\
+\textbf{Credit   Approval}      & 0.885                                 & 0.885 $\pm$ 0.19\%                                 & {\color[HTML]{FE0000} \textbf{0.889 $\pm$ 0.15\%}} & 0.885 $\pm$ 0.17\%                                 & 0.886 $\pm$ 0.20\%                                 & 0.888 $\pm$ 0.13\%                                 \\
+\textbf{Digits}                 & {\color[HTML]{FE0000} \textbf{0.984}} & 0.983 $\pm$ 0.14\%                                 & \textbf{0.984 $\pm$ 0.06\%}                        & 0.982 $\pm$ 0.06\%                                 & 0.983 $\pm$ 0.09\%                                 & \textbf{0.984 $\pm$ 0.07\%}                        \\
+\textbf{Ecoli}                  & 0.848                                 & 0.857 $\pm$ 0.42\%                                 & \textbf{0.861 $\pm$ 0.53\%}                        & 0.859 $\pm$ 0.40\%                                 & 0.857 $\pm$ 0.53\%                                 & {\color[HTML]{FE0000} \textbf{0.861 $\pm$ 0.41\%}} \\
+\textbf{Parkinsons}             & {\color[HTML]{FE0000} 0.963}          & 0.961 $\pm$ 0.28\%                                 & 0.965 $\pm$ 0.27\%                                 & 0.959 $\pm$ 0.32\%                                 & 0.962 $\pm$ 0.20\%                                 & \textbf{0.964 $\pm$ 0.30\%}                        \\
+\textbf{Semeion}                & {\color[HTML]{FE0000} 0.945}          & 0.941 $\pm$ 0.14\%                                 & \textbf{0.946 $\pm$ 0.23\%}                        & 0.940 $\pm$ 0.11\%                                 & 0.943 $\pm$ 0.28\%                                 & 0.945 $\pm$ 0.15\%                                 \\
+\textbf{Statlog   Segmentation} & 0.976                                 & 0.977 $\pm$ 0.12\%                                 & \textbf{0.979 $\pm$ 0.17\%}                        & 0.976 $\pm$ 0.18\%                                 & 0.978 $\pm$ 0.04\%                                 & {\color[HTML]{FE0000} \textbf{0.979 $\pm$ 0.04\%}} \\
+\textbf{Wilt}                   & {\color[HTML]{FE0000} 0.906}          & 0.904 $\pm$ 0.15\%                                 & \textbf{0.907 $\pm$ 0.26\%}                        & 0.906 $\pm$ 0.23\%                                 & 0.906 $\pm$ 0.03\%                                 & \textbf{0.907 $\pm$ 0.14\%}                        \\
+\textbf{Zoo}                    & {\color[HTML]{FE0000} \textbf{1.000}} & 0.988 $\pm$ 0.97\%                                 & 0.990 $\pm$ 0.90\%                                 & 0.995 $\pm$ 0.80\%                                 & 0.996 $\pm$ 0.86\%                                 & 0.995 $\pm$ 0.81\%                                 \\ \hline
+\end{tabular}%
+}
+\end{table}
+
 \begin{table}[ht]
@@
 \centering
-\includegraphics[width=0.85\textwidth]{Fig8}
+\includegraphics[width=\textwidth]{fig_time}
 \caption{Time for tuning hyperparameters of the SVC algorithm to obtain good values of quality metrics on the Wilt dataset} \label{fig8}

### 2023-10-02 2023/MMST/samplepaper.tex
fix item 5 in modification discription
fix reviewer comments

@@
 
-Global optimization deals with the development of methods for solving problems of finding the global (absolute) minima of multidimensional multiextremal functions.
+Global optimization deals with the development of methods for solving problems of finding the global (absolute) minimum of multidimensional multiextremal functions.
 The possibility of constructing a reliable estimate of the global extremum in such problems is principally based on the available information on the problem to be solved.
@@
 \begin{equation}\label{eq2} 
-\phi^*=\phi(y(x^* ))=\min_{x \in [0,1]} \left\{ \phi(y(x)) \right\},
+f^*(x)=\phi(y(x^* ))=\min_{x \in [0,1]} \left\{ \phi(y(x)) \right\},
 \end{equation}
@@
 \begin{equation}\label{eq4} 
-| \phi (x')-\phi (x'') | \leq K \| x'-x'' \| ^{1/N}, \; x',x'' \in [0,1],
+| \phi (x')-\phi (x'') | \leq K \rho(x',x''), \; x',x'' \in [0,1],
 \end{equation}
-where $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the relation $K \leq 2L\sqrt {N+3}$ \cite{Strongin2000}.
+where $\rho(x',x'') =  |x' - x''|^{1/N}$ and $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the relation $K \leq 2L\sqrt {N+3}$ \cite{Strongin2000}.
 
@@
 The first trial will be executed in the middle point $x^1 \in (0,1)$.
-The choice of the point $x^{k+1}, k \leq 1$ of the next $(k+1)^\text{th}$ trial is performed on the base of the following rules.
+The choice of the point $x^{k+1}, k \geq 1$ of the next $(k+1)^\text{th}$ trial is performed on the base of the following rules.
 
-\paragraph{Rule 1.} Renumber (by the lower indices) the points $x^i, 1 \leq i \leq k$  of the previous trials in the ascending order by their coordinates, i.e.
+\paragraph{Rule 1.} Renumber (by the lower indices) the points $x^i, 0 \leq i \leq k$  of the previous trials in the ascending order by their coordinates, i.e.
 \begin{equation}\label{eq5} 
-0=x_0 < x_1 < ... < x_i < ... < x_k=1
+0=x_0 < x_1 < ... < x_i < ... < x_{k}=1
 \end{equation}
-and juxtapose them with the values $z_i=\phi (x_i), 1 \leq i \leq k$, computed at these points and to the indices $v_i=v(x_i)$ defined by the rule
+and juxtapose them with the values $z_i=\phi (x_i), 0 < i < k$, computed at these points and to the indices $v_i=v(x_i)$ defined by the rule
 \begin{equation}\label{eq6} 
@@
 \end{equation}
-The points $x_0=0$ and $x_{k+1}=1$ are introduced additionally (the values $z_0$ and $z_{k+1}$ are undefined) for convenience of the following notations.
+The points $x_0=0$ and $x_{k}=1$ are introduced additionally (the values $z_0$ and $z_{k}$ are undefined) for convenience of the following notations.
 
@@
 
-\paragraph{Rule 4.} For each interval $(x_{i-1},x_i),1 \leq i \leq k+1$, compute the value $R(i)$ called a \textit{characteristic} of the interval according to the formula
+\paragraph{Rule 4.} For each interval $(x_{i-1},x_i),1 \leq i \leq k$, compute the value $R(i)$ called a \textit{characteristic} of the interval according to the formula
 \begin{equation}\label{eq9} 
@@
 \[
-R(t)= \max\{R(i): \; 1 \leq i \leq k+1\}.
+R(t)= \max\{R(i): \; 1 \leq i \leq k\}.
 \]
@@
 
-  \item For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$, the computing of the characteristic $R(i)$ is performed according to the following formulae.
+  \item For each interval $(x_{i-1}, x_i), 1 \leq i \leq k$, the computing of the characteristic $R(i)$ is performed according to the following formulae.
 
@@
 \end{equation}
-  \item If several intervals have equal characteristics, prefer the interval with computable boundary points that has the lowest number.
+  \item If several intervals have equal characteristics, prefer the interval with computable points that has the lowest number.
   \item If the best interval appears to be the one with two non-computable points, perform the next trial in the middle point of the interval.

### 2023-10-03 2023/MMST/samplepaper.tex
replace one reference

@@
 
-A number of efficient deterministic algorithms for solving the Lipschitz global optimization problems have been developed ~\cite{Grishagin2018,Jones2021,PaulaviciusZilinskas2014,Birect2020,Sergeyev2017}.
+A number of efficient deterministic algorithms for solving the Lipschitz global optimization problems have been developed ~\cite{Grishagin2016_2,Jones2021,PaulaviciusZilinskas2014,Birect2020,Sergeyev2017}.
 The comparisons conducted have shown the deterministic algorithms outperform (in various criteria) the widespread nature-inspired algorithms ~\cite{Liberti2005,Sergeyev2018}.

### 2023-10-03 2023/ITMO paper/elsarticle-template-num.tex
first version of the scheme
time table

@@
 Next, we consider the performance of the iOpt, Optuna, and Hyperopt frameworks.
-Table \ref{tab:3} shows the running time of the frameworks for optimizing the hyperparameters of the SVC algorithm.
+Table \ref{tab:time} shows the running time of the frameworks for optimizing the hyperparameters.
 A single optimization run was performed with a maximum number of 200 search trials.
@@
 
-\begin{table}[ht]
-\caption{Time of hyperparameter tuning for a single run of algorithms.}
-\label{tab:3}
-\begin{tabular}{lllllll}
+% Please add the following required packages to your document preamble:
+% \usepackage{graphicx}
+\begin{table}[]
+\centering
+\caption{Time of hyperparameter tuning.}
+\label{tab:time}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{cccccc}
 \hline
-Dataset &
-  \begin{tabular}[c]{@{}l@{}}Optuna \\ random\end{tabular} &
-  \begin{tabular}[c]{@{}l@{}}Optuna \\ tpe\end{tabular} &
-  \begin{tabular}[c]{@{}l@{}}Optuna \\ cmaes\end{tabular} &
-  \begin{tabular}[c]{@{}l@{}}Optuna \\ nsgaii\end{tabular} &
-  \begin{tabular}[c]{@{}l@{}}Hyperopt\\ tpe\end{tabular} &
-  iOpt \\ \hline
-Balance                                                         & 10.8   & 9.2    & 9.0    & 10.6   & 10.3   & \textbf{7.7}    \\
-\begin{tabular}[c]{@{}l@{}}Bank \\ Marketing\end{tabular}       & 2888.5 & 3406.5 & \textbf{2704.7} & 3078.6 & 3190.3 & 2963.3 \\
-Banknote                                                        & 31.9   & \textbf{12.0}   & 17.7   & 22.1   & 21.7   & 18.4   \\
-\begin{tabular}[c]{@{}l@{}}Breast \\ Cancer\end{tabular}        & 9.3    & 7.6    & 6.3    & 8.1    & 8.5    & \textbf{6.0}    \\
-CarEvaluation                                                   & 50.9   & 63.3   & 51.0   & 53.1   & 69.5   & \textbf{45.2}   \\
-CNAE9                                                           & 295.6  & 207.6  & 318.4  & 262.2  & 253.5  & \textbf{144.9}  \\
-\begin{tabular}[c]{@{}l@{}}Credit \\ Approval\end{tabular}      & 14.2   & 12.6   & 13.4   & 13.4   & 15.8   & \textbf{10.8}   \\
-Digits                                                          & 132.8  & \textbf{57.5}   & 82.1   & 111.6  & 92.4   & 62.5   \\
-Ecoli                                                           & 5.5    & 6.3    & 4.0    & 5.4    & 6.4    & \textbf{3.6}    \\
-Parkinsons                                                      & 3.5    & 5.0    & 3.6    & 3.8    & 4.4    & \textbf{2.9}    \\
-Semeion                                                         & 218.4  & 119.5  & 157.9  & 135.7  & 193.6  & \textbf{114.2}  \\
-\begin{tabular}[c]{@{}l@{}}Statlog \\ Segmentation\end{tabular} & 156.1  & \textbf{55.4}   & 87.1   & 159.2  & 91.6   & 66.5   \\
-Wilt                                                            & 70.0   & 75.8   & 68.5   & 67.4   & 78.2   & \textbf{60.9}   \\
-Zoo                                                             & 3.0    & 3.9    & 3.0    & 3.1    & 4.2    & \textbf{1.8}    \\ \hline
-\end{tabular}
+\textbf{Dataset}                & \textbf{iOpt}     & \textbf{Optuna(tpe)}        & \textbf{Optuna(cmaes)} & \textbf{Optuna(nsgaii)}    & \textbf{Hyperopt(tpe)}      \\ \hline
+                                & \multicolumn{5}{c}{\textbf{SVM}}                                                                                                    \\ \hline
+\textbf{Balance}                & \textbf{00:00:15} & 00:00:19 ± 0.66\%           & 00:00:17 ± 8.71\%      & 00:00:20 ± 3.00\%          & 00:00:21 ± 4.44\%           \\
+\textbf{Bank-marketing}         & \textbf{01:31:18} & 02:02:05 ± 4.12\%           & 01:49:42 ± 8.35\%      & 01:51:06 ± 4.62\%          & 01:52:03 ± 8.25\%           \\
+\textbf{Banknote}               & 00:00:35          & \textbf{00:00:25 ± 3.60\%}  & 00:00:32 ± 10.86\%     & 00:00:48 ± 5.89\%          & 00:00:39 ± 4.37\%           \\
+\textbf{Breast Cancer}          & \textbf{00:00:11} & 00:00:15 ± 1.38\%           & 00:00:14 ± 3.45\%      & 00:00:16 ± 4.59\%          & 00:00:18 ± 3.79\%           \\
+\textbf{CarEvaluation}          & \textbf{00:01:21} & 00:01:48 ± 2.82\%           & 00:01:28 ± 6.69\%      & 00:01:34 ± 2.01\%          & 00:01:44 ± 3.10\%           \\
+\textbf{CNAE-9}                 & \textbf{00:05:53} & 00:08:24 ± 18.64\%          & 00:08:30 ± 3.92\%      & 00:08:27 ± 2.28\%          & 00:08:22 ± 5.14\%           \\
+\textbf{Credit   Approval}      & \textbf{00:00:19} & 00:00:26 ± 16.46\%          & 00:00:21 ± 7.39\%      & 00:00:23 ± 2.38\%          & 00:00:25 ± 4.47\%           \\
+\textbf{Digits}                 & 00:02:07          & \textbf{00:02:01 ± 5.04\%}  & 00:02:47 ± 7.83\%      & 00:03:35 ± 7.10\%          & 00:03:09 ± 4.74\%           \\
+\textbf{Ecoli}                  & \textbf{00:00:08} & 00:00:13 ± 2.09\%           & 00:00:10 ± 2.28\%      & 00:00:10 ± 1.66\%          & 00:00:14 ± 3.30\%           \\
+\textbf{Parkinsons}             & \textbf{00:00:07} & 00:00:11 ± 1.38\%           & 00:00:08 ± 0.96\%      & 00:00:08 ± 1.43\%          & 00:00:12 ± 4.09\%           \\
+\textbf{Semeion}                & \textbf{00:04:03} & 00:04:16 ± 5.22\%           & 00:05:17 ± 3.24\%      & 00:05:51 ± 1.73\%          & 00:05:57 ± 3.06\%           \\
+\textbf{Statlog   Segmentation} & 00:02:07          & \textbf{00:01:45 ± 4.02\%}  & 00:02:59 ± 5.75\%      & 00:04:12 ± 6.43\%          & 00:03:05 ± 4.06\%           \\
+\textbf{Wilt}                   & \textbf{00:01:51} & 00:02:17 ± 6.65\%           & 00:02:01 ± 9.83\%      & 00:02:07 ± 3.28\%          & 00:02:16 ± 4.78\%           \\
+\textbf{Zoo}                    & \textbf{00:00:05} & 00:00:10 ± 2.08\%           & 00:00:07 ± 1.73\%      & 00:00:07 ± 2.42\%          & 00:00:10 ± 3.47\%           \\ \hline
+                                & \multicolumn{5}{c}{\textbf{XGBoost}}                                                                                                \\ \hline
+\textbf{Balance}                & \textbf{00:01:49} & 00:03:38 ± 10.57\%          & 00:03:09 ± 10.94\%     & 00:02:36 ± 6.39\%          & 00:02:55 ± 11.63\%          \\
+\textbf{Bank-marketing}         & 01:46:27          & 01:52:51 ± 17.65\%          & 02:06:44 ± 14.76\%     & 01:51:28 ± 7.79\%          & \textbf{01:43:15 ± 13.30\%} \\
+\textbf{Banknote}               & \textbf{00:02:02} & 00:03:36 ± 18.51\%          & 00:03:15 ± 10.33\%     & 00:02:35 ± 6.93\%          & 00:02:45 ± 5.20\%           \\
+\textbf{Breast Cancer}          & \textbf{00:01:45} & 00:03:25 ± 26.50\%          & 00:02:33 ± 19.95\%     & 00:02:31 ± 13.55\%         & 00:02:26 ± 8.54\%           \\
+\textbf{CarEvaluation}          & 00:10:10          & 00:12:05 ± 8.62\%           & 00:10:53 ± 8.07\%      & 00:07:14 ± 6.79\%          & \textbf{00:07:42 ± 12.51\%} \\
+\textbf{CNAE-9}                 & 03:47:42          & 04:55:03 ± 26.76\%          & 05:43:27 ± 14.35\%     & 03:55:49 ± 13.82\%         & \textbf{03:42:39 ± 18.30\%} \\
+\textbf{Credit   Approval}      & \textbf{00:01:30} & 00:02:27 ± 23.63\%          & 00:01:54 ± 23.08\%     & 00:01:58 ± 6.65\%          & 00:01:44 ± 13.34\%          \\
+\textbf{Digits}                 & 00:48:57          & 00:59:09 ± 21.83\%          & 00:54:27 ± 8.20\%      & 00:43:35 ± 9.93\%          & \textbf{00:46:52 ± 8.24\%}  \\
+\textbf{Ecoli}                  & \textbf{00:02:40} & 00:04:03 ± 21.05\%          & 00:03:30 ± 22.08\%     & 00:02:48 ± 8.63\%          & 00:03:05 ± 8.97\%           \\
+\textbf{Parkinsons}             & \textbf{00:01:09} & 00:02:32 ± 16.25\%          & 00:01:52 ± 12.30\%     & 00:01:34 ± 4.24\%          & 00:01:45 ± 12.30\%          \\
+\textbf{Semeion}                & 01:52:08          & 01:19:57 ± 12.21\%          & 01:42:52 ± 14.47\%     & 02:01:16 ± 9.22\%          & \textbf{01:41:40 ± 7.68\%}  \\
+\textbf{Statlog   Segmentation} & 00:34:59          & 00:43:59 ± 13.68\%          & 00:38:13 ± 6.14\%      & \textbf{00:25:03 ± 8.20\%} & 00:32:15 ± 8.61\%           \\
+\textbf{Wilt}                   & 00:06:44          & 00:10:42 ± 6.89\%           & 00:08:20 ± 7.94\%      & \textbf{00:06:11 ± 8.76\%} & 00:08:09 ± 8.43\%           \\
+\textbf{Zoo}                    & \textbf{00:01:47} & 00:02:53 ± 20.08\%          & 00:02:21 ± 16.16\%     & 00:02:05 ± 4.28\%          & 00:02:13 ± 9.09\%           \\ \hline
+                                & \multicolumn{5}{c}{\textbf{MLP}}                                                                                                    \\ \hline
+\textbf{Balance}                & \textbf{00:09:29} & 00:18:14 ± 61.97\%          & 00:26:45 ± 11.25\%     & 00:20:55 ± 11.15\%         & 00:15:37 ± 10.06\%          \\
+\textbf{Bank-marketing}         & 06:10:05          & \textbf{04:42:08 ± 39.48\%} & 05:44:30 ± 15.70\%     & 07:04:15 ± 10.63\%         & 08:11:02 ± 16.22\%          \\
+\textbf{Banknote}               & \textbf{02:07:36} & 02:29:16 ± 44.57\%          & 04:43:15 ± 13.11\%     & 03:20:56 ± 8.58\%          & 02:39:20 ± 11.03\%          \\
+\textbf{Breast Cancer}          & \textbf{00:43:14} & 02:48:26 ± 10.35\%          & 02:41:29 ± 15.72\%     & 02:46:45 ± 5.13\%          & 02:11:56 ± 29.01\%          \\
+\textbf{CarEvaluation}          & \textbf{02:08:22} & 04:21:15 ± 17.08\%          & 03:04:09 ± 12.40\%     & 03:45:18 ± 9.32\%          & 03:55:44 ± 19.04\%          \\
+\textbf{CNAE-9}                 & 07:46:15          & \textbf{07:26:02 ± 9.33\%}  & 13:29:53 ± 11.28\%     & 10:57:27 ± 12.39\%         & 08:32:12 ± 7.19\%           \\
+\textbf{Credit   Approval}      & \textbf{00:43:32} & 02:41:30 ± 10.89\%          & 02:24:56 ± 15.16\%     & 02:39:29 ± 4.92\%          & 01:57:22 ± 29.27\%          \\
+\textbf{Digits}                 & \textbf{02:58:50} & 10:50:47 ± 7.21\%           & 09:30:39 ± 5.32\%      & 10:19:20 ± 6.95\%          & 08:45:23 ± 29.04\%          \\
+\textbf{Ecoli}                  & \textbf{00:05:52} & 00:08:21 ± 16.93\%          & 00:07:03 ± 16.70\%     & 00:08:08 ± 7.34\%          & 00:07:01 ± 12.44\%          \\
+\textbf{Parkinsons}             & \textbf{00:16:21} & 00:46:32 ± 34.33\%          & 00:49:54 ± 14.73\%     & 00:44:24 ± 5.05\%          & 00:34:48 ± 14.89\%          \\
+\textbf{Semeion}                & \textbf{03:38:23} & 11:26:28 ± 25.20\%          & 10:19:44 ± 7.97\%      & 09:01:10 ± 13.98\%         & 07:55:17 ± 20.55\%          \\
+\textbf{Statlog   Segmentation} & \textbf{00:48:57} & 19:27:00 ± 15.47\%          & 17:09:16 ± 11.57\%     & 17:02:03 ± 3.17\%          & 11:20:02 ± 3.89\%           \\
+\textbf{Wilt}                   & \textbf{01:23:04} & 17:57:58 ± 28.49\%          & 19:23:37 ± 11.57\%     & 22:06:21 ± 9.34\%          & 12:38:43 ± 3.14\%           \\
+\textbf{Zoo}                    & 00:01:46          & 00:01:52 ± 22.83\%          & 00:01:58 ± 7.72\%      & \textbf{00:01:40 ± 8.30\%} & 00:02:01 ± 12.96\%          \\ \hline
+\end{tabular}%
+}
 \end{table}
+
+%\begin{table}[ht]
+%\caption{Time of hyperparameter tuning for a single run of algorithms.}
+%\label{tab:3}
+%\begin{tabular}{lllllll}
+%\hline
+%Dataset &
+%  \begin{tabular}[c]{@{}l@{}}Optuna \\ random\end{tabular} &
+%  \begin{tabular}[c]{@{}l@{}}Optuna \\ tpe\end{tabular} &
+%  \begin{tabular}[c]{@{}l@{}}Optuna \\ cmaes\end{tabular} &
+%  \begin{tabular}[c]{@{}l@{}}Optuna \\ nsgaii\end{tabular} &
+%  \begin{tabular}[c]{@{}l@{}}Hyperopt\\ tpe\end{tabular} &
+%  iOpt \\ \hline
+%Balance                                                         & 10.8   & 9.2    & 9.0    & 10.6   & 10.3   & \textbf{7.7}    \\
+%\begin{tabular}[c]{@{}l@{}}Bank \\ Marketing\end{tabular}       & 2888.5 & 3406.5 & \textbf{2704.7} & 3078.6 & 3190.3 & 2963.3 \\
+%Banknote                                                        & 31.9   & \textbf{12.0}   & 17.7   & 22.1   & 21.7   & 18.4   \\
+%\begin{tabular}[c]{@{}l@{}}Breast \\ Cancer\end{tabular}        & 9.3    & 7.6    & 6.3    & 8.1    & 8.5    & \textbf{6.0}    \\
+%CarEvaluation                                                   & 50.9   & 63.3   & 51.0   & 53.1   & 69.5   & \textbf{45.2}   \\
+%CNAE9                                                           & 295.6  & 207.6  & 318.4  & 262.2  & 253.5  & \textbf{144.9}  \\
+%\begin{tabular}[c]{@{}l@{}}Credit \\ Approval\end{tabular}      & 14.2   & 12.6   & 13.4   & 13.4   & 15.8   & \textbf{10.8}   \\
+%Digits                                                          & 132.8  & \textbf{57.5}   & 82.1   & 111.6  & 92.4   & 62.5   \\
+%Ecoli                                                           & 5.5    & 6.3    & 4.0    & 5.4    & 6.4    & \textbf{3.6}    \\
+%Parkinsons                                                      & 3.5    & 5.0    & 3.6    & 3.8    & 4.4    & \textbf{2.9}    \\
+%Semeion                                                         & 218.4  & 119.5  & 157.9  & 135.7  & 193.6  & \textbf{114.2}  \\
+%\begin{tabular}[c]{@{}l@{}}Statlog \\ Segmentation\end{tabular} & 156.1  & \textbf{55.4}   & 87.1   & 159.2  & 91.6   & 66.5   \\
+%Wilt                                                            & 70.0   & 75.8   & 68.5   & 67.4   & 78.2   & \textbf{60.9}   \\
+%Zoo                                                             & 3.0    & 3.9    & 3.0    & 3.1    & 4.2    & \textbf{1.8}    \\ \hline
+%\end{tabular}
+%\end{table}
 

### 2023-10-03 2023/ITMO paper/elsarticle-template-num.tex
fix text

@@
 The framework is open source and is available at \url{https://github.com/aimclub/iOpt}.
+\textcolor[rgb]{1,0,0}{The general structure of the framework is shown in Fig. \ref{fig_iOpt}. 
+Description of the main algorithms used in iOpt is given in the following sections.
+}
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\textwidth]{scheme.pdf}
+\caption{iOpt general structure} \label{fig_iOpt}
+\end{figure}
 
@@
 \label{tab:1}
-\begin{tabular}{lcccccc}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{ccccccc}
 \hline
 Name                                                            & Num   & Attributes & Int & Float & Categorical & Classes \\ \hline
-Balance                                                         & 625   & 4          & 4   & 0     & 0           & 3       \\
-Bank Marketing                                                  & 45211 & 20         & 6   & 4     & 10          & 2       \\
-Banknote                                                        & 1372  & 4          & 0   & 4     & 0           & 2       \\
-Breast Cancer                                                   & 569   & 9          & 9   & 0     & 0           & 2       \\
-CarEvaluation                                                   & 1728  & 6          & 0   & 0     & 6           & 4       \\
-CNAE9                                                           & 1080  & 856        & 856 & 0     & 0           & 9       \\
-Credit Approval                                                 & 690   & 15         & 0   & 6     & 9           & 2       \\
-Digits                                                          & 5620  & 64         & 64  & 0     & 0           & 10      \\
-Ecoli                                                           & 336   & 7          & 0   & 7     & 0           & 8       \\
-Parkinsons                                                      & 197   & 22         & 0   & 22    & 0           & 2       \\
-Semeion                                                         & 1593  & 256        & 256 & 0     & 0           & 10      \\
-\begin{tabular}[c]{@{}l@{}}Statlog \\ Segmentation\end{tabular} & 2310  & 19         & 1   & 18    & 0           & 7       \\
-Wilt                                                            & 4889  & 5          & 0   & 5     & 0           & 2       \\
-Zoo                                                             & 101   & 16         & 16  & 0     & 0           & 7       \\ \hline
+\textbf{Balance}                                                         & 625   & 4          & 4   & 0     & 0           & 3       \\
+\textbf{Bank Marketing}                                                  & 45211 & 20         & 6   & 4     & 10          & 2       \\
+\textbf{Banknote}                                                        & 1372  & 4          & 0   & 4     & 0           & 2       \\
+\textbf{Breast Cancer}                                                   & 569   & 9          & 9   & 0     & 0           & 2       \\
+\textbf{CarEvaluation}                                                   & 1728  & 6          & 0   & 0     & 6           & 4       \\
+\textbf{CNAE-9}                                                           & 1080  & 856        & 856 & 0     & 0           & 9       \\
+\textbf{Credit Approval}                                                 & 690   & 15         & 0   & 6     & 9           & 2       \\
+\textbf{Digits}                                                          & 5620  & 64         & 64  & 0     & 0           & 10      \\
+\textbf{Ecoli}                                                           & 336   & 7          & 0   & 7     & 0           & 8       \\
+\textbf{Parkinsons}                                                      & 197   & 22         & 0   & 22    & 0           & 2       \\
+\textbf{Semeion}                                                         & 1593  & 256        & 256 & 0     & 0           & 10      \\
+\textbf{Statlog Segmentation}                                            & 2310  & 19         & 1   & 18    & 0           & 7       \\
+\textbf{Wilt}                                                            & 4889  & 5          & 0   & 5     & 0           & 2       \\
+\textbf{Zoo}                                                             & 101   & 16         & 16  & 0     & 0           & 7       \\ \hline
 \end{tabular}
+}
 \end{table}
@@
 \label{tab:2}
-\begin{tabular}{lccc}
+\begin{tabular}{cccc}
 \hline
@@
 When using the iOpt framework, we applied a variant of computation in which the available resource was divided into portions by the number of trials.
-Most of the trials (75\%) were performed using the global search algorithm (see Section \ref{sec_iOpt}), while in the remaining trials we used the local method to refine the solution found \cite{Kelley}.
+Most of the trials (85\%) were performed using the global search algorithm (see Section \ref{sec_iOpt}), while in the remaining trials we used the local method to refine the solution found \cite{Kelley}.
 
@@
 
-Computational experiments were performed on the computing nodes of the Lobachevsky supercomputer (Intel Xeon Silver 4310T CPU @ 2.30GHz, 64Gb RAM, Linux CentOS 7).
+Computational experiments were performed on the computing nodes of the Lobachevsky supercomputer (Intel Xeon CPU E5-2660 @ 2.20GHz, 64Gb RAM, Linux CentOS 7).
 We used the following versions of frameworks for the experiments: iOpt 0.2.22, Optuna 3.1.1, Hyperopt 0.2.7.
@@
 
-Figures \ref{fig_res_1} and \ref{fig_res_2} show the values of the classification metrics of the SVC algorithm in the worst and average case, respectively.
+The values of the classification metrics and its standard deviation of the SVC, XGBoost and MLP algorithms are presented in the table \ref{tab:res}.
+The table bold shows the best values for averaging the results.
+Values in case of unsuccessful launch of stochastic algorithms are highlighted in red.
+
+It is easy to see from the table \ref{tab:res} that our iOpt framework is comparable in terms of quality to Hyperopt and Optuna.
+When using various methods for constructing classification models and different datasets based on the average metric values, iOpt allows us to obtain the best values in 15 models out of 42.
+It is especially worth noting that the algorithms implemented in both Optuna and Hyperopt contain random parameters; as a consequence, the best values of the objective metric are not guaranteed for a single run.
+To obtain a guaranteed result, the stochastic algorithm must be run several times.
+Looking at the unsuccessful launch Optuna or Hyperopt framework, our framework shows better results in more than half of the cases (23 out of 42 models).
+In most of the remaining models, the value of quality metrics lags only slightly.
+iOpt shows the best results when tuning hyperparameters for training the SVC model.
+
+%Figures \ref{fig_res_1} and \ref{fig_res_2} show the values of the classification metrics of the SVC algorithm in the worst and average case, respectively.
 As can be seen, when optimizing the hyperparameters of the SVC algorithm on the Banknote, CNAE-9, Statlog Segmentation, and Wilt datasets, the iOpt framework has demonstrated the best results.
 
-\begin{figure}
-\centering
-\includegraphics[width=0.85\textwidth]{SVC_1.pdf}
-\caption{Values of the optimized metrics for the SVC algorithm, obtained in the worst case} \label{fig_res_1}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.85\textwidth]{SVC_2_avg.pdf}
-\caption{Values of optimized metrics for the SVC algorithm averaged over 10 runs} \label{fig_res_2}
-\end{figure}
-
-On the CarEvaluation, Credit Approval, Ecoli, Parkinsons, and Semeion datasets, the best values of the metrics on average were obtained with the Optuna (tpe).
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.85\textwidth]{SVC_1.pdf}
+%\caption{Values of the optimized metrics for the SVC algorithm, obtained in the worst case} \label{fig_res_1}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.85\textwidth]{SVC_2_avg.pdf}
+%\caption{Values of optimized metrics for the SVC algorithm averaged over 10 runs} \label{fig_res_2}
+%\end{figure}
+%
+%
+%On the CarEvaluation, Credit Approval, Ecoli, Parkinsons, and Semeion datasets, the best values of the metrics on average were obtained with the Optuna (tpe).
 On the datasets Balance, Breast Cancer, Digits, on average the Optuna(cmaes) performed well.
@@
 In the case of the Zoo dataset, all optimization algorithms showed the same results.
-
-It is especially worth noting that the algorithms implemented in both Optuna and Hyperopt contain random parameters; as a consequence, the best values of the objective metric are not guaranteed for a single run.
-
-This fact is illustrated by Fig.
+%
+%It is especially worth noting that the algorithms implemented in both Optuna and Hyperopt contain random parameters; as a consequence, the best values of the objective metric are not guaranteed for a single run.
+%
+%This fact is illustrated by Fig.
 \ref{fig_res_1}.
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.85\textwidth]{XGBoost_1.pdf}
-\caption{Values of the optimized metrics for the XGBoost algorithm, obtained in the worst case} \label{fig_res_3}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.85\textwidth]{XGBoost_2_avg.pdf}
-\caption{Values of optimized metrics for the XGBoost algorithm averaged over 10 runs} \label{fig_res_4}
-\end{figure}
-
-Figures \ref{fig_res_3} and \ref{fig_res_4} show the value of classification metrics of the XGBoost algorithm in the worst and average case, respectively.
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.85\textwidth]{XGBoost_1.pdf}
+%\caption{Values of the optimized metrics for the XGBoost algorithm, obtained in the worst case} \label{fig_res_3}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.85\textwidth]{XGBoost_2_avg.pdf}
+%\caption{Values of optimized metrics for the XGBoost algorithm averaged over 10 runs} \label{fig_res_4}
+%\end{figure}
+%
+%Figures \ref{fig_res_3} and \ref{fig_res_4} show the value of classification metrics of the XGBoost algorithm in the worst and average case, respectively.
 In the case of optimizing the hyperparameters  of the XGBoost algorithm, the iOpt framework showed better results on the CarEvaluation, CNAE-9, Credit Approval, Ecoli, Parkinsons, Wilt, Zoo datasets.
-
-On average, when using 10 runs, the best performance of the Optuna(tpe) framework was observed on the Balance, Bank-marketing, Banknote, Breast Cancer, CarEvaluation, Credit Approval, Digits, Ecoli, Semeion, Statlog Segmentation, Wilt datasets.
+%
+%On average, when using 10 runs, the best performance of the Optuna(tpe) framework was observed on the Balance, Bank-marketing, Banknote, Breast Cancer, CarEvaluation, Credit Approval, Digits, Ecoli, Semeion, Statlog Segmentation, Wilt datasets.
 The Hyperopt(tpe) showed good results on the CNAE-9 and Parkinsons datasets on average for 10 runs.
 
-\begin{figure}
-\centering
-\includegraphics[width=0.85\textwidth]{MLP_1.pdf}
-\caption{Values of the optimized metrics for the XGBoost algorithm, obtained in the worst case} \label{fig_res_5}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.85\textwidth]{MLP_2_avg.pdf}
-\caption{Values of optimized metrics for the XGBoost algorithm averaged over 10 runs} \label{fig_res_6}
-\end{figure}
-
-Figures \ref{fig_res_5} and \ref{fig_res_6} show the values of classification metrics of the MLP algorithm.
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.85\textwidth]{MLP_1.pdf}
+%\caption{Values of the optimized metrics for the XGBoost algorithm, obtained in the worst case} \label{fig_res_5}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.85\textwidth]{MLP_2_avg.pdf}
+%\caption{Values of optimized metrics for the XGBoost algorithm averaged over 10 runs} \label{fig_res_6}
+%\end{figure}
+%
+%Figures \ref{fig_res_5} and \ref{fig_res_6} show the values of classification metrics of the MLP algorithm.
 The iOpt framework performed better on the Balance, Bank-marketing, Banknote, Breast Cancer, Credit Approval, Wilt, Zoo datasets.
 Considering the average values of metrics on 10 runs of hyperparameter optimization, the Optuna framework performed better on the Breast Cancer, CNAE-9, Credit Approval, Digits, Ecoli, Parkinsons, Semeion datasets, and the Hyperopt on the Statlog Segmentation, Wilt, CarEvaluation datasets.
+%
 
-Thus, we can conclude that the iOpt framework shows comparable results with the Optuna and Hyperopt frameworks in terms of performance quality.
+%Thus, we can conclude that the iOpt framework shows comparable results with the Optuna and Hyperopt frameworks in terms of performance quality.
 It should be reminded that the algorithms implemented within the iOpt framework are deterministic.
 As a consequence, iOpt, unlike Optuna and Hyperopt, only needs to be run once to achieve the best values of the quality metric.
-
-Next, we consider the performance of the iOpt, Optuna, and Hyperopt frameworks.
-Table \ref{tab:time} shows the running time of the frameworks for optimizing the hyperparameters.
-A single optimization run was performed with a maximum number of 200 search trials.
-It is obvious that even in the case of a single run iOpt shows better performance on many datasets compared to Optuna and Hyperopt.
-
-To obtain a good solution, the randomized methods from Optuna and Hyperopt should be repeated several times.
-In this case, the performance gain of iOpt is even greater.
-As an example, Figure \ref{fig8} shows a comparison of the runtime of the frameworks with 10 reruns of Optuna and Hyperopt on the Wilt dataset.
 
@@
 % If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
-\begin{table}[]
+\begin{table}[ht]
 \centering
@@
 \end{table}
+
+Next, we consider the performance of the iOpt, Optuna, and Hyperopt frameworks.
+Table \ref{tab:time} shows the running time of the frameworks for optimizing the hyperparameters\footnote{Since the random search did not show good results for the optimized metrics, the columns were excluded.}.
+A single optimization run was also performed with a maximum number of 200 search trials.
+As before, the best results when considering the average time value are shown in bold.
+It is obvious that even in the case of a single run iOpt shows better performance on many datasets compared to Optuna and Hyperopt (29 out of 42 models).
+
+To obtain a good solution, the stochastic methods from Optuna and Hyperopt should be repeated several times.
+In this case, the performance gain of iOpt is even greater.
+As an example, Figure \ref{fig8} shows a comparison of the hyperparameter tuning times of iOpt, Optuna and Hyperopt on several datasets and different classification models.
+For iOpt, the time is shown as a blue dot.
+For Optuna and Hyperopt, the average computation time is displayed, as well as the spread of values.
 
@@
 \includegraphics[width=\textwidth]{fig_time}
-\caption{Time for tuning hyperparameters of the SVC algorithm to obtain good values of quality metrics on the Wilt dataset} \label{fig8}
+\caption{Time of hyperparameter tuning} \label{fig8}
 \end{figure}

### 2023-10-04 2023/MMST/samplepaper.tex
last fix after first review

@@
 \begin{equation}\label{eq4} 
-| \phi (x')-\phi (x'') | \leq K \rho(x',x''), \; x',x'' \in [0,1],
+| f(x')-f(x'') | \leq K \rho(x',x''), \; x',x'' \in [0,1],
 \end{equation}
@@
 \end{equation}
-and juxtapose them with the values $z_i=\phi (x_i), 0 < i < k$, computed at these points and to the indices $v_i=v(x_i)$ defined by the rule
+and juxtapose them with the values $z_i=f(x_i), 0 < i < k$, computed at these points and to the indices $v_i=v(x_i)$ defined by the rule
 \begin{equation}\label{eq6} 
@@
 \end{equation}
-of the unknown H{\"o}lder constant $K$ from (\ref{eq4}) for the reduced function $\phi(x)$, where $r>1$ is a parameter of the algorithm; if $\mu=0$ accept $\mu=1$.
+of the unknown H{\"o}lder constant $K$ from (\ref{eq4}) for the reduced function $f(x)$, where $r>1$ is a parameter of the algorithm; if $\mu=0$ accept $\mu=1$.
 \paragraph{Rule 3.} Determine the best current value of the objective function
 \begin{equation}\label{eq8} 
-z^*=\min \left\{ \phi (x_i): 1\leq i \leq k \right\}.
+z^*=\min \left\{ f(x_i): 1\leq i \leq k \right\}.
 \end{equation}
@@
     -\varepsilon _r, & {\quad v=0 ,}\\
-    \min \left\{ \phi (x_i): v(x_i)=1, 1 \leq i \leq k \right\}, & {\quad v=1 ,}
+    \min \left\{ f(x_i): v(x_i)=1, 0 < i < k \right\}, & {\quad v=1 ,}
   \end{cases}

### 2023-10-06 2023/MMST/samplepaper.tex
add one more reference

@@
 In such subregions, one cannot perform the numerical modeling and compute the objective function values correctly.
-This phenomenon can be interpreted either as the presence of some hidden constraints \cite{Stripinis2021} or as a partial computability of the objective function in the search domain \cite{Candelieri2019,Sergeyev2003,Strongin2020}.
+This phenomenon can be interpreted either as the presence of some hidden constraints \cite{Stripinis2021}, or as unknown discontinuous regions \cite{Audet2022}, or as a partial computability of the objective function in the search domain \cite{Candelieri2019,Sergeyev2003,Strongin2020}.
 In such a statement, the optimization problem becomes more complex since the region of feasible parameters' combinations is not defined in advance.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{fig3.png}
+\centering
+\includegraphics[width=0.95\textwidth]{fig3.png}
 \caption{Imputation of the non-computable values by the averages of the neighbor points. Approach No. 3} \label{fig3}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{fig4.png}
+\centering
+\includegraphics[width=0.95\textwidth]{fig4.png}
 \caption{Imputation of non-computable values by averages of the neighboring points. Approach No. 4} \label{fig4}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{fig5.png}
+\centering
+\includegraphics[width=0.95\textwidth]{fig5.png}
 \caption{Imputation of non-computable values by averages of the minorant and the majorant. Approach No. 5} \label{fig5}

### 2023-10-08 2023/ITMO paper/elsarticle-template-num.tex
fix translation

@@
 \affiliation[UNN]{organization={Lobachevsky University},%Department and Organization
-            addressline={Gagarin Av. 23}, 
+            addressline={Gagarin Ave. 23}, 
             city={Nizhni Novgorod},
@@
 \includegraphics[width=0.8\textwidth]{Fig1.pdf}
-\caption{The values of the function $f(x)$ must be inside the area highlighted by colour} \label{fig1}
+\caption{The values of the function $f(x)$ must be inside the area highlighted by colour.} \label{fig1}
 \end{figure}
@@
 The framework is open source and is available at \url{https://github.com/aimclub/iOpt}.
-\textcolor[rgb]{1,0,0}{The general structure of the framework is shown in Fig. \ref{fig_iOpt}. 
-Description of the main algorithms used in iOpt is given in the following sections.
-}
+The general structure of the framework is shown in Fig.~\ref{fig_iOpt}. The main algorithms used in iOpt are described in the following sections.
+
 \begin{figure}[t]
@@
 \includegraphics[width=1.0\textwidth]{scheme.pdf}
-\caption{iOpt general structure} \label{fig_iOpt}
+\caption{iOpt general structure.} \label{fig_iOpt}
 \end{figure}
@@
 
-The values of the classification metrics and its standard deviation of the SVC, XGBoost and MLP algorithms are presented in the table \ref{tab:res}.
-The table bold shows the best values for averaging the results.
-Values in case of unsuccessful launch of stochastic algorithms are highlighted in red.
+In order to check the statistical significance of the experimental results, the Friedman test was applied.
+An implementation of the corresponding test from the SciPy library was used.
+The target metric values obtained by Optuna, Hyperopt and iOpt framework runs were used as input data.
+The test was carried out separately for each of the algorithms for which hyperparameters were tuned: SVC, XGBoost, MLP.
+The null hypothesis about the consistent work of the compared methods is rejected at significance level $p=0.01$.
+Note that in some cases the $p$-value was undefined (for example, when analyzing the ``Banknote'' data set using the MLP method and ``Zoo'' using the SVC method), which is due to the equality of the target metric for all optimization runs.
 
-It is easy to see from the table \ref{tab:res} that our iOpt framework is comparable in terms of quality to Hyperopt and Optuna.
+The values of the classification metrics and their standard deviation for the SVC, XGBoost and MLP algorithms are presented in Table \ref{tab:res}. 
+The values corresponding to the best result on average are highlighted in bold. The values corresponding to the guaranteed best result, i.e. the best result for the worst case, are highlighted in red.
+
+It is obvious from Table \ref{tab:res} that the iOpt framework is comparable in terms of quality to Hyperopt and Optuna.
 When using various methods for constructing classification models and different datasets based on the average metric values, iOpt allows us to obtain the best values in 15 models out of 42.
-It is especially worth noting that the algorithms implemented in both Optuna and Hyperopt contain random parameters; as a consequence, the best values of the objective metric are not guaranteed for a single run.
-To obtain a guaranteed result, the stochastic algorithm must be run several times.
-Looking at the unsuccessful launch Optuna or Hyperopt framework, our framework shows better results in more than half of the cases (23 out of 42 models).
-In most of the remaining models, the value of quality metrics lags only slightly.
-iOpt shows the best results when tuning hyperparameters for training the SVC model.
+It is worth noting that algorithms which involve random variables in one way or another do not guarantee the best values of the objective function for one run.
+To obtain good results, a stochastic algorithm must be run several times.
+If one focuses on a guaranteed result, iOpt works better in more than half of the cases (23 out of 42 models).
+iOpt shows the best results when tuning hyperparameters for the SVC model.
+For most of the remaining models, the value of quality metrics differs only slightly.
 
@@
 % If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
-\begin{table}[ht]
+\begin{table}
 \centering
-\caption{Values of optimized metrics for the algorithm's averaged over 10 runs}
+\caption{Values of optimized metrics for the algorithms averaged over 10 runs}
 \label{tab:res}
@@
 
-Next, we consider the performance of the iOpt, Optuna, and Hyperopt frameworks.
-Table \ref{tab:time} shows the running time of the frameworks for optimizing the hyperparameters\footnote{Since the random search did not show good results for the optimized metrics, the columns were excluded.}.
-A single optimization run was also performed with a maximum number of 200 search trials.
-As before, the best results when considering the average time value are shown in bold.
-It is obvious that even in the case of a single run iOpt shows better performance on many datasets compared to Optuna and Hyperopt (29 out of 42 models).
+Next, we consider the performance of the iOpt, Optuna, and Hyperopt frameworks. Table \ref{tab:time} shows the running time of the frameworks for tuning the hyperparameters.
+Since the random search did not show good results in terms of the evaluation metrics, we did not include the corresponding column in the table. 
+A single optimization run was performed with a maximum of 200 search trials.
+The values corresponding to the best result on average are highlighted in bold.
+It can be clearly seen that even in the case of a single run iOpt shows better performance on many datasets compared to Optuna and Hyperopt (29 out of 42 models).
 
-To obtain a good solution, the stochastic methods from Optuna and Hyperopt should be repeated several times.
-In this case, the performance gain of iOpt is even greater.
-As an example, Figure \ref{fig8} shows a comparison of the hyperparameter tuning times of iOpt, Optuna and Hyperopt on several datasets and different classification models.
+To obtain a good solution, the methods from Optuna and Hyperopt should be restarted several times.
+If one focuses on the total operating time, the performance gain of iOpt is even greater.
+As an example, Figure \ref{fig8} shows a comparison of the hyperparameter tuning times for iOpt, Optuna and Hyperopt on several datasets and different classification models.
 For iOpt, the time is shown as a blue dot.
-For Optuna and Hyperopt, the average computation time is displayed, as well as the spread of values.
+For Optuna and Hyperopt, the box plots are shown.
 
@@
 % \usepackage{graphicx}
-\begin{table}[]
+\begin{table}
 \centering
@@
                                 & \multicolumn{5}{c}{\textbf{SVM}}                                                                                                    \\ \hline
-\textbf{Balance}                & \textbf{00:00:15} & 00:00:19 ± 0.66\%           & 00:00:17 ± 8.71\%      & 00:00:20 ± 3.00\%          & 00:00:21 ± 4.44\%           \\
-\textbf{Bank-marketing}         & \textbf{01:31:18} & 02:02:05 ± 4.12\%           & 01:49:42 ± 8.35\%      & 01:51:06 ± 4.62\%          & 01:52:03 ± 8.25\%           \\
-\textbf{Banknote}               & 00:00:35          & \textbf{00:00:25 ± 3.60\%}  & 00:00:32 ± 10.86\%     & 00:00:48 ± 5.89\%          & 00:00:39 ± 4.37\%           \\
-\textbf{Breast Cancer}          & \textbf{00:00:11} & 00:00:15 ± 1.38\%           & 00:00:14 ± 3.45\%      & 00:00:16 ± 4.59\%          & 00:00:18 ± 3.79\%           \\
-\textbf{CarEvaluation}          & \textbf{00:01:21} & 00:01:48 ± 2.82\%           & 00:01:28 ± 6.69\%      & 00:01:34 ± 2.01\%          & 00:01:44 ± 3.10\%           \\
-\textbf{CNAE-9}                 & \textbf{00:05:53} & 00:08:24 ± 18.64\%          & 00:08:30 ± 3.92\%      & 00:08:27 ± 2.28\%          & 00:08:22 ± 5.14\%           \\
-\textbf{Credit   Approval}      & \textbf{00:00:19} & 00:00:26 ± 16.46\%          & 00:00:21 ± 7.39\%      & 00:00:23 ± 2.38\%          & 00:00:25 ± 4.47\%           \\
-\textbf{Digits}                 & 00:02:07          & \textbf{00:02:01 ± 5.04\%}  & 00:02:47 ± 7.83\%      & 00:03:35 ± 7.10\%          & 00:03:09 ± 4.74\%           \\
-\textbf{Ecoli}                  & \textbf{00:00:08} & 00:00:13 ± 2.09\%           & 00:00:10 ± 2.28\%      & 00:00:10 ± 1.66\%          & 00:00:14 ± 3.30\%           \\
-\textbf{Parkinsons}             & \textbf{00:00:07} & 00:00:11 ± 1.38\%           & 00:00:08 ± 0.96\%      & 00:00:08 ± 1.43\%          & 00:00:12 ± 4.09\%           \\
-\textbf{Semeion}                & \textbf{00:04:03} & 00:04:16 ± 5.22\%           & 00:05:17 ± 3.24\%      & 00:05:51 ± 1.73\%          & 00:05:57 ± 3.06\%           \\
-\textbf{Statlog   Segmentation} & 00:02:07          & \textbf{00:01:45 ± 4.02\%}  & 00:02:59 ± 5.75\%      & 00:04:12 ± 6.43\%          & 00:03:05 ± 4.06\%           \\
-\textbf{Wilt}                   & \textbf{00:01:51} & 00:02:17 ± 6.65\%           & 00:02:01 ± 9.83\%      & 00:02:07 ± 3.28\%          & 00:02:16 ± 4.78\%           \\
-\textbf{Zoo}                    & \textbf{00:00:05} & 00:00:10 ± 2.08\%           & 00:00:07 ± 1.73\%      & 00:00:07 ± 2.42\%          & 00:00:10 ± 3.47\%           \\ \hline
+\textbf{Balance}                & \textbf{00:00:15} & 00:00:19 $\pm$ 0.66\%           & 00:00:17 $\pm$ 8.71\%      & 00:00:20 $\pm$ 3.00\%          & 00:00:21 $\pm$ 4.44\%           \\
+\textbf{Bank-marketing}         & \textbf{01:31:18} & 02:02:05 $\pm$ 4.12\%           & 01:49:42 $\pm$ 8.35\%      & 01:51:06 $\pm$ 4.62\%          & 01:52:03 $\pm$ 8.25\%           \\
+\textbf{Banknote}               & 00:00:35          & \textbf{00:00:25 $\pm$ 3.60\%}  & 00:00:32 $\pm$ 10.86\%     & 00:00:48 $\pm$ 5.89\%          & 00:00:39 $\pm$ 4.37\%           \\
+\textbf{Breast Cancer}          & \textbf{00:00:11} & 00:00:15 $\pm$ 1.38\%           & 00:00:14 $\pm$ 3.45\%      & 00:00:16 $\pm$ 4.59\%          & 00:00:18 $\pm$ 3.79\%           \\
+\textbf{CarEvaluation}          & \textbf{00:01:21} & 00:01:48 $\pm$ 2.82\%           & 00:01:28 $\pm$ 6.69\%      & 00:01:34 $\pm$ 2.01\%          & 00:01:44 $\pm$ 3.10\%           \\
+\textbf{CNAE-9}                 & \textbf{00:05:53} & 00:08:24 $\pm$ 18.64\%          & 00:08:30 $\pm$ 3.92\%      & 00:08:27 $\pm$ 2.28\%          & 00:08:22 $\pm$ 5.14\%           \\
+\textbf{Credit   Approval}      & \textbf{00:00:19} & 00:00:26 $\pm$ 16.46\%          & 00:00:21 $\pm$ 7.39\%      & 00:00:23 $\pm$ 2.38\%          & 00:00:25 $\pm$ 4.47\%           \\
+\textbf{Digits}                 & 00:02:07          & \textbf{00:02:01 $\pm$ 5.04\%}  & 00:02:47 $\pm$ 7.83\%      & 00:03:35 $\pm$ 7.10\%          & 00:03:09 $\pm$ 4.74\%           \\
+\textbf{Ecoli}                  & \textbf{00:00:08} & 00:00:13 $\pm$ 2.09\%           & 00:00:10 $\pm$ 2.28\%      & 00:00:10 $\pm$ 1.66\%          & 00:00:14 $\pm$ 3.30\%           \\
+\textbf{Parkinsons}             & \textbf{00:00:07} & 00:00:11 $\pm$ 1.38\%           & 00:00:08 $\pm$ 0.96\%      & 00:00:08 $\pm$ 1.43\%          & 00:00:12 $\pm$ 4.09\%           \\
+\textbf{Semeion}                & \textbf{00:04:03} & 00:04:16 $\pm$ 5.22\%           & 00:05:17 $\pm$ 3.24\%      & 00:05:51 $\pm$ 1.73\%          & 00:05:57 $\pm$ 3.06\%           \\
+\textbf{Statlog   Segmentation} & 00:02:07          & \textbf{00:01:45 $\pm$ 4.02\%}  & 00:02:59 $\pm$ 5.75\%      & 00:04:12 $\pm$ 6.43\%          & 00:03:05 $\pm$ 4.06\%           \\
+\textbf{Wilt}                   & \textbf{00:01:51} & 00:02:17 $\pm$ 6.65\%           & 00:02:01 $\pm$ 9.83\%      & 00:02:07 $\pm$ 3.28\%          & 00:02:16 $\pm$ 4.78\%           \\
+\textbf{Zoo}                    & \textbf{00:00:05} & 00:00:10 $\pm$ 2.08\%           & 00:00:07 $\pm$ 1.73\%      & 00:00:07 $\pm$ 2.42\%          & 00:00:10 $\pm$ 3.47\%           \\ \hline
                                 & \multicolumn{5}{c}{\textbf{XGBoost}}                                                                                                \\ \hline
-\textbf{Balance}                & \textbf{00:01:49} & 00:03:38 ± 10.57\%          & 00:03:09 ± 10.94\%     & 00:02:36 ± 6.39\%          & 00:02:55 ± 11.63\%          \\
-\textbf{Bank-marketing}         & 01:46:27          & 01:52:51 ± 17.65\%          & 02:06:44 ± 14.76\%     & 01:51:28 ± 7.79\%          & \textbf{01:43:15 ± 13.30\%} \\
-\textbf{Banknote}               & \textbf{00:02:02} & 00:03:36 ± 18.51\%          & 00:03:15 ± 10.33\%     & 00:02:35 ± 6.93\%          & 00:02:45 ± 5.20\%           \\
-\textbf{Breast Cancer}          & \textbf{00:01:45} & 00:03:25 ± 26.50\%          & 00:02:33 ± 19.95\%     & 00:02:31 ± 13.55\%         & 00:02:26 ± 8.54\%           \\
-\textbf{CarEvaluation}          & 00:10:10          & 00:12:05 ± 8.62\%           & 00:10:53 ± 8.07\%      & 00:07:14 ± 6.79\%          & \textbf{00:07:42 ± 12.51\%} \\
-\textbf{CNAE-9}                 & 03:47:42          & 04:55:03 ± 26.76\%          & 05:43:27 ± 14.35\%     & 03:55:49 ± 13.82\%         & \textbf{03:42:39 ± 18.30\%} \\
-\textbf{Credit   Approval}      & \textbf{00:01:30} & 00:02:27 ± 23.63\%          & 00:01:54 ± 23.08\%     & 00:01:58 ± 6.65\%          & 00:01:44 ± 13.34\%          \\
-\textbf{Digits}                 & 00:48:57          & 00:59:09 ± 21.83\%          & 00:54:27 ± 8.20\%      & 00:43:35 ± 9.93\%          & \textbf{00:46:52 ± 8.24\%}  \\
-\textbf{Ecoli}                  & \textbf{00:02:40} & 00:04:03 ± 21.05\%          & 00:03:30 ± 22.08\%     & 00:02:48 ± 8.63\%          & 00:03:05 ± 8.97\%           \\
-\textbf{Parkinsons}             & \textbf{00:01:09} & 00:02:32 ± 16.25\%          & 00:01:52 ± 12.30\%     & 00:01:34 ± 4.24\%          & 00:01:45 ± 12.30\%          \\
-\textbf{Semeion}                & 01:52:08          & 01:19:57 ± 12.21\%          & 01:42:52 ± 14.47\%     & 02:01:16 ± 9.22\%          & \textbf{01:41:40 ± 7.68\%}  \\
-\textbf{Statlog   Segmentation} & 00:34:59          & 00:43:59 ± 13.68\%          & 00:38:13 ± 6.14\%      & \textbf{00:25:03 ± 8.20\%} & 00:32:15 ± 8.61\%           \\
-\textbf{Wilt}                   & 00:06:44          & 00:10:42 ± 6.89\%           & 00:08:20 ± 7.94\%      & \textbf{00:06:11 ± 8.76\%} & 00:08:09 ± 8.43\%           \\
-\textbf{Zoo}                    & \textbf{00:01:47} & 00:02:53 ± 20.08\%          & 00:02:21 ± 16.16\%     & 00:02:05 ± 4.28\%          & 00:02:13 ± 9.09\%           \\ \hline
+\textbf{Balance}                & \textbf{00:01:49} & 00:03:38 $\pm$ 10.57\%          & 00:03:09 $\pm$ 10.94\%     & 00:02:36 $\pm$ 6.39\%          & 00:02:55 $\pm$ 11.63\%          \\
+\textbf{Bank-marketing}         & 01:46:27          & 01:52:51 $\pm$ 17.65\%          & 02:06:44 $\pm$ 14.76\%     & 01:51:28 $\pm$ 7.79\%          & \textbf{01:43:15 $\pm$ 13.30\%} \\
+\textbf{Banknote}               & \textbf{00:02:02} & 00:03:36 $\pm$ 18.51\%          & 00:03:15 $\pm$ 10.33\%     & 00:02:35 $\pm$ 6.93\%          & 00:02:45 $\pm$ 5.20\%           \\
+\textbf{Breast Cancer}          & \textbf{00:01:45} & 00:03:25 $\pm$ 26.50\%          & 00:02:33 $\pm$ 19.95\%     & 00:02:31 $\pm$ 13.55\%         & 00:02:26 $\pm$ 8.54\%           \\
+\textbf{CarEvaluation}          & 00:10:10          & 00:12:05 $\pm$ 8.62\%           & 00:10:53 $\pm$ 8.07\%      & 00:07:14 $\pm$ 6.79\%          & \textbf{00:07:42 $\pm$ 12.51\%} \\
+\textbf{CNAE-9}                 & 03:47:42          & 04:55:03 $\pm$ 26.76\%          & 05:43:27 $\pm$ 14.35\%     & 03:55:49 $\pm$ 13.82\%         & \textbf{03:42:39 $\pm$ 18.30\%} \\
+\textbf{Credit   Approval}      & \textbf{00:01:30} & 00:02:27 $\pm$ 23.63\%          & 00:01:54 $\pm$ 23.08\%     & 00:01:58 $\pm$ 6.65\%          & 00:01:44 $\pm$ 13.34\%          \\
+\textbf{Digits}                 & 00:48:57          & 00:59:09 $\pm$ 21.83\%          & 00:54:27 $\pm$ 8.20\%      & 00:43:35 $\pm$ 9.93\%          & \textbf{00:46:52 $\pm$ 8.24\%}  \\
+\textbf{Ecoli}                  & \textbf{00:02:40} & 00:04:03 $\pm$ 21.05\%          & 00:03:30 $\pm$ 22.08\%     & 00:02:48 $\pm$ 8.63\%          & 00:03:05 $\pm$ 8.97\%           \\
+\textbf{Parkinsons}             & \textbf{00:01:09} & 00:02:32 $\pm$ 16.25\%          & 00:01:52 $\pm$ 12.30\%     & 00:01:34 $\pm$ 4.24\%          & 00:01:45 $\pm$ 12.30\%          \\
+\textbf{Semeion}                & 01:52:08          & 01:19:57 $\pm$ 12.21\%          & 01:42:52 $\pm$ 14.47\%     & 02:01:16 $\pm$ 9.22\%          & \textbf{01:41:40 $\pm$ 7.68\%}  \\
+\textbf{Statlog   Segmentation} & 00:34:59          & 00:43:59 $\pm$ 13.68\%          & 00:38:13 $\pm$ 6.14\%      & \textbf{00:25:03 $\pm$ 8.20\%} & 00:32:15 $\pm$ 8.61\%           \\
+\textbf{Wilt}                   & 00:06:44          & 00:10:42 $\pm$ 6.89\%           & 00:08:20 $\pm$ 7.94\%      & \textbf{00:06:11 $\pm$ 8.76\%} & 00:08:09 $\pm$ 8.43\%           \\
+\textbf{Zoo}                    & \textbf{00:01:47} & 00:02:53 $\pm$ 20.08\%          & 00:02:21 $\pm$ 16.16\%     & 00:02:05 $\pm$ 4.28\%          & 00:02:13 $\pm$ 9.09\%           \\ \hline
                                 & \multicolumn{5}{c}{\textbf{MLP}}                                                                                                    \\ \hline
-\textbf{Balance}                & \textbf{00:09:29} & 00:18:14 ± 61.97\%          & 00:26:45 ± 11.25\%     & 00:20:55 ± 11.15\%         & 00:15:37 ± 10.06\%          \\
-\textbf{Bank-marketing}         & 06:10:05          & \textbf{04:42:08 ± 39.48\%} & 05:44:30 ± 15.70\%     & 07:04:15 ± 10.63\%         & 08:11:02 ± 16.22\%          \\
-\textbf{Banknote}               & \textbf{02:07:36} & 02:29:16 ± 44.57\%          & 04:43:15 ± 13.11\%     & 03:20:56 ± 8.58\%          & 02:39:20 ± 11.03\%          \\
-\textbf{Breast Cancer}          & \textbf{00:43:14} & 02:48:26 ± 10.35\%          & 02:41:29 ± 15.72\%     & 02:46:45 ± 5.13\%          & 02:11:56 ± 29.01\%          \\
-\textbf{CarEvaluation}          & \textbf{02:08:22} & 04:21:15 ± 17.08\%          & 03:04:09 ± 12.40\%     & 03:45:18 ± 9.32\%          & 03:55:44 ± 19.04\%          \\
-\textbf{CNAE-9}                 & 07:46:15          & \textbf{07:26:02 ± 9.33\%}  & 13:29:53 ± 11.28\%     & 10:57:27 ± 12.39\%         & 08:32:12 ± 7.19\%           \\
-\textbf{Credit   Approval}      & \textbf{00:43:32} & 02:41:30 ± 10.89\%          & 02:24:56 ± 15.16\%     & 02:39:29 ± 4.92\%          & 01:57:22 ± 29.27\%          \\
-\textbf{Digits}                 & \textbf{02:58:50} & 10:50:47 ± 7.21\%           & 09:30:39 ± 5.32\%      & 10:19:20 ± 6.95\%          & 08:45:23 ± 29.04\%          \\
-\textbf{Ecoli}                  & \textbf{00:05:52} & 00:08:21 ± 16.93\%          & 00:07:03 ± 16.70\%     & 00:08:08 ± 7.34\%          & 00:07:01 ± 12.44\%          \\
-\textbf{Parkinsons}             & \textbf{00:16:21} & 00:46:32 ± 34.33\%          & 00:49:54 ± 14.73\%     & 00:44:24 ± 5.05\%          & 00:34:48 ± 14.89\%          \\
-\textbf{Semeion}                & \textbf{03:38:23} & 11:26:28 ± 25.20\%          & 10:19:44 ± 7.97\%      & 09:01:10 ± 13.98\%         & 07:55:17 ± 20.55\%          \\
-\textbf{Statlog   Segmentation} & \textbf{00:48:57} & 19:27:00 ± 15.47\%          & 17:09:16 ± 11.57\%     & 17:02:03 ± 3.17\%          & 11:20:02 ± 3.89\%           \\
-\textbf{Wilt}                   & \textbf{01:23:04} & 17:57:58 ± 28.49\%          & 19:23:37 ± 11.57\%     & 22:06:21 ± 9.34\%          & 12:38:43 ± 3.14\%           \\
-\textbf{Zoo}                    & 00:01:46          & 00:01:52 ± 22.83\%          & 00:01:58 ± 7.72\%      & \textbf{00:01:40 ± 8.30\%} & 00:02:01 ± 12.96\%          \\ \hline
+\textbf{Balance}                & \textbf{00:09:29} & 00:18:14 $\pm$ 61.97\%          & 00:26:45 $\pm$ 11.25\%     & 00:20:55 $\pm$ 11.15\%         & 00:15:37 $\pm$ 10.06\%          \\
+\textbf{Bank-marketing}         & 06:10:05          & \textbf{04:42:08 $\pm$ 39.48\%} & 05:44:30 $\pm$ 15.70\%     & 07:04:15 $\pm$ 10.63\%         & 08:11:02 $\pm$ 16.22\%          \\
+\textbf{Banknote}               & \textbf{02:07:36} & 02:29:16 $\pm$ 44.57\%          & 04:43:15 $\pm$ 13.11\%     & 03:20:56 $\pm$ 8.58\%          & 02:39:20 $\pm$ 11.03\%          \\
+\textbf{Breast Cancer}          & \textbf{00:43:14} & 02:48:26 $\pm$ 10.35\%          & 02:41:29 $\pm$ 15.72\%     & 02:46:45 $\pm$ 5.13\%          & 02:11:56 $\pm$ 29.01\%          \\
+\textbf{CarEvaluation}          & \textbf{02:08:22} & 04:21:15 $\pm$ 17.08\%          & 03:04:09 $\pm$ 12.40\%     & 03:45:18 $\pm$ 9.32\%          & 03:55:44 $\pm$ 19.04\%          \\
+\textbf{CNAE-9}                 & 07:46:15          & \textbf{07:26:02 $\pm$ 9.33\%}  & 13:29:53 $\pm$ 11.28\%     & 10:57:27 $\pm$ 12.39\%         & 08:32:12 $\pm$ 7.19\%           \\
+\textbf{Credit   Approval}      & \textbf{00:43:32} & 02:41:30 $\pm$ 10.89\%          & 02:24:56 $\pm$ 15.16\%     & 02:39:29 $\pm$ 4.92\%          & 01:57:22 $\pm$ 29.27\%          \\
+\textbf{Digits}                 & \textbf{02:58:50} & 10:50:47 $\pm$ 7.21\%           & 09:30:39 $\pm$ 5.32\%      & 10:19:20 $\pm$ 6.95\%          & 08:45:23 $\pm$ 29.04\%          \\
+\textbf{Ecoli}                  & \textbf{00:05:52} & 00:08:21 $\pm$ 16.93\%          & 00:07:03 $\pm$ 16.70\%     & 00:08:08 $\pm$ 7.34\%          & 00:07:01 $\pm$ 12.44\%          \\
+\textbf{Parkinsons}             & \textbf{00:16:21} & 00:46:32 $\pm$ 34.33\%          & 00:49:54 $\pm$ 14.73\%     & 00:44:24 $\pm$ 5.05\%          & 00:34:48 $\pm$ 14.89\%          \\
+\textbf{Semeion}                & \textbf{03:38:23} & 11:26:28 $\pm$ 25.20\%          & 10:19:44 $\pm$ 7.97\%      & 09:01:10 $\pm$ 13.98\%         & 07:55:17 $\pm$ 20.55\%          \\
+\textbf{Statlog   Segmentation} & \textbf{00:48:57} & 19:27:00 $\pm$ 15.47\%          & 17:09:16 $\pm$ 11.57\%     & 17:02:03 $\pm$ 3.17\%          & 11:20:02 $\pm$ 3.89\%           \\
+\textbf{Wilt}                   & \textbf{01:23:04} & 17:57:58 $\pm$ 28.49\%          & 19:23:37 $\pm$ 11.57\%     & 22:06:21 $\pm$ 9.34\%          & 12:38:43 $\pm$ 3.14\%           \\
+\textbf{Zoo}                    & 00:01:46          & 00:01:52 $\pm$ 22.83\%          & 00:01:58 $\pm$ 7.72\%      & \textbf{00:01:40 $\pm$ 8.30\%} & 00:02:01 $\pm$ 12.96\%          \\ \hline
 \end{tabular}%
@@
 \includegraphics[width=\textwidth]{fig_time}
-\caption{Time of hyperparameter tuning} \label{fig8}
+\caption{Time of hyperparameter tuning.} \label{fig8}
 \end{figure}

### 2023-10-14 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
fix commas

@@
 	\begin{abstract}
-Decision making models described as problems of multicriterial optimization are very complicated for investigation because the property of criteria contradictoriness leads to the notion of the solution as the set of non-dominated parameters (Pareto set).
+Decision making models described as problems of multicriterial optimization are very complicated for investigation, because the property of criteria contradictoriness leads to the notion of the solution as the set of non-dominated parameters (Pareto set).
 The complexity of these problems increases significantly in the case where criteria are multiextremal.
@@
 As a core of novelty, this algorithm includes in its computational scheme a machine learning procedure with combination of accumulating the information of solved subproblems that allows one to significantly accelerate building the Pareto set.
-Effectiveness of the proposed approach is estimated in the representative computational experiment on test sets of multicriterial problems with multiextremal criteria for different dimensions in comparison with several nature-inspired multiobjective optimization algorithms.
+Effectiveness of the proposed approach is estimated in the representative computational experiment on test sets of multicriterial problems with multiextremal criteria for different dimensions, in comparison with several nature-inspired multiobjective optimization algorithms.
 
@@
 
-The third approach is based on the reduction of multicriterial problem to a family of single-criterion (scalar) optimization subproblems by means of convolution method and solving these subproblems by efficient mathematical programming algorithms \cite{Collette2004,Marler2009,Marler2004,Pardalos2017,Wierzbicki,Ehrgott2005,GergelKozinov2020}.
+The third approach is based on the reduction of multicriterial problem to a family of single-criterion (scalar) optimization subproblems by means of convolution method, and solving these subproblems by efficient mathematical programming algorithms \cite{Collette2004,Marler2009,Marler2004,Pardalos2017,Wierzbicki,Ehrgott2005,GergelKozinov2020}.
 Under corresponding conditions, the solution of the scalar subproblem is a Pareto-optimal point.
@@
 Further, a special global optimization algorithm designed on the ISA platform is applied for solving the univariate problems.
-Its computational scheme includes two accelerators that decreases significantly number of algorithm's iterations.
+Its computational scheme includes two accelerators that significantly decrease the number of algorithm's iterations.
 The first one when solving current one-dimensional problem takes into account results of iterations of all the previous problems solved already.
@@
 
-Unfortunately, finding a decision   that provides the minimal values for all partial criteria simultaneously is possible only in exceptional cases because, as a rule, partial criteria are contradictory, when decreasing one criterion leads to increasing the other.
+Unfortunately, finding a decision   that provides the minimal values for all partial criteria simultaneously is possible only in exceptional cases because, as a rule, partial criteria may be contradictory, i.e.
+when decreasing one criterion leads to increasing the other.
 
@@
 
-If in the domain $H$ all the criteria 
+If in the domain $H$ all the criteria satisfy
 \begin{equation}
@@
 
-However, because of fundamental characteristics of the mapping $y(x)$ the reduced function $g(y(x))$ does not retain Lipschitz property in the case of Lipschitzian partial criteria from (\ref{eq:02}).
+However, because of fundamental characteristics of the mapping $y(x)$, the reduced function $g(y(x))$ does not retain Lipschitz property in the case of Lipschitzian partial criteria from (\ref{eq:02}).
 In this situation the function $g(y(x))$ satisfies the H{\" o}lder condition
@@
     \label{eq:12}
-    \mu^i=(\mu^i_k, \dots, \mu^i_k), \; 1 \leq i \leq m.
+    \mu^i=(\mu^i_1, \dots, \mu^i_k), \; 1 \leq i \leq m.
 \end{equation}
@@
 
-It should be noted that if the value (\ref{eq:17}) has been already calculated and the vector $\Omega^i(x^j)$ has been stored, then evaluation of another function with its weight coefficient set $M'=(\mu'^1, \dots, \mu'^k)$ at the point $x^j$ does not require the recalculation the partial criteria (\ref{eq:02}) because the value $z^{l, j}$ can be easily computed according to (\ref{eq:17}).
+It should be noted that if the value (\ref{eq:17}) has been already calculated and the vector $\Omega^i(x^j)$ has been stored, then evaluation of another function with its weight coefficient set $M'=(\mu'^1, \dots, \mu'^k)$ at the point $x^j$ does not require the recalculation of the partial criteria (\ref{eq:02}) because the value $z^{l, j}$ can be easily computed according to (\ref{eq:17}).
 
@@
 In all the stages after the first one $(i>1)$, the trial coordinates obtained in all previous stages are taken as points of initial trials at current stage.
-As the matrix $A^{i-1}_q$ contains values of all partial criteria at these points it is easy to recalculate values $z^{i,1}, z^{i,2}, \dots, z^{i,p}$ of function $f^i(x)$ according to (\ref{eq:17}) and to build the updated matrix $A^i_q$.
+As the matrix $A^{i-1}_q$ contains values of all partial criteria at these points, it is easy to recalculate values $z^{i,1}, z^{i,2}, \dots, z^{i,p}$ of function $f^i(x)$ according to (\ref{eq:17}) and to build the updated matrix $A^i_q$.
 
@@
 
-\textbf{Rule 4.} In the criteria space on the base of the marked up matrix $A^i_p$ a separating hyperplane $L_p$ is constructed applying logistic regression \cite{Yu2011} as efficient tool of machine learning.
+\textbf{Rule 4.} In the criteria space on the base of the marked up matrix $A^i_p$, a separating hyperplane $L_p$ is constructed applying logistic regression \cite{Yu2011} as efficient tool of machine learning.
 
@@
 
-The data of the table show that AGS-SP and AGS-R solve the problem (\ref{eq:28}) more efficient than evolutionary methods because while spending the same number of trials (and even less as AGS-SP) they provide the  better HV index and find more points in Pareto set approximation.
+The data of the table shows that AGS-SP and AGS-R solve the problem (\ref{eq:28}) more efficiently than evolutionary methods because while spending the same number of trials (and even less as AGS-SP) they provide the  better HV index and find more points in Pareto set approximation.
 If to compare AGS-SP and AGS-R with each other, then AGS-SP has an advantage due to HV indices being the same, but AGS-SP was ``faster'' as it executed fewer trials.

### 2023-10-16 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
fix NUMTA paper

@@
 
-As the decision-making model we consider the multicriterial multidimensional optimization problems with multiextremal ``black-box'' criteria herein after called as MMCO problems.
+As the decision-making model we study the multicriterial multidimensional optimization problems with multiextremal ``black-box'' criteria hereinafter called as MMCO problems.
+The property of multiextremality means that any criterion can have several local minima.
 Combination of multicriteriality and multiextremality generates the dramatic complexity of these problems, because, first, criteria are contradictory, as a rule, and this circumstance leads to a complicated concept of solution (Pareto set) in such problems and, second, in the multiextremal case the growth of computational efforts is exponential when increasing the dimensionality (number of independent variables).
@@
 In the class of deterministic methods for multiextremal optimization with guaranteed convergence to global optimum \cite{Gablonsky2001,Sergeyev2010,Sergeyev2006,Evtushenko2014,Strongin2003,Sergeyev2013}, the algorithms developed in the framework of the information-statistical approach (ISA) \cite{Gergel2018,Barkalov2018} are among the most effective.
+In the ISA framework, the optimized function is considered as a sample of a random function with predefined stochastic properties (information), and the optimization method is designed as a statistical procedure that optimizes the probability of global minimizer finding.
 The present paper describes a new algorithm belonging to the ISA class and aimed at solving MMCO problems that accelerates the Pareto set building by the application of machine learning ideas.
@@
 
-Unfortunately, finding a decision   that provides the minimal values for all partial criteria simultaneously is possible only in exceptional cases because, as a rule, partial criteria may be contradictory, i.e.
+Unfortunately, finding a decision that provides the global minimal values for all partial criteria simultaneously is possible only in exceptional cases because, as a rule, partial criteria may be contradictory, i.e.
 when decreasing one criterion leads to increasing the other.
@@
 
-In general case, the functions satisfying the Lipschitz condition are multiextremal, therefore, the convolution (\ref{eq:05}) will be multiextremal and Lipschitzian as well.
+In should be noted, that under the conditions (\ref{eq:09}), the convolution (\ref{eq:05}) is Lipschitzian as well.
+Moreover, this convolution is multiextremal in the case of multiextremality of partial criteria $\varphi_s(y)$, $1 \leq s \leq k$.
 It means that for solving the problem (\ref{eq:08}) it is required to apply methods of nonlinear programming which provide finding the global optimum.
@@
 The following computational scheme is proposed for solving the problems of the family (\ref{eq:15}).
-In this scheme, all the problems (\ref{eq:15}) are solved sequentially and during minimization of current function $f^i(x), i > 1$, the results of preceding problems $f^j(x) \to \min, 1 \leq j \leq i - 1,$ completed already are taken into account.
+In this scheme, all the problems (\ref{eq:15}) are solved sequentially and during minimization of current function $f^i(x), i > 1$, the results of preceding problems $f^j(x) \to \min, 1 \leq j \leq i - 1$, completed already are taken into account.
+The main idea is to calculate values of the function $f_i(x)$ by utilizing the values of partial criteria computed earlier and to incorporate them to the scheme of the global search algorithm as results of initial iterations.
 
@@
 Three series of experiments were executed.
-The first one dealt with a simple problem and was aimed at demonstration of positive influence of criteria space analysis by means of separating planes and at comparing with evolutionary techniques.
+The first one dealt with a simple problem and was aimed at efficiency demonstration of the proposed technique that combines the ISA method with separating planes (AGS-SP) compared to its prototype AGS-R and evolutionary algorithms.
 This problem was bi-criteria one \cite{CHIANDUSSI2012912}:

### 2023-10-16 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
fix numta

@@
 The Hyper-Volume (HV) index reflecting the quality of the Pareto set approximation was used as a comparison criterion.
-The results of the experiment have demonstrated significant advantages of the new algorithm over other methods having involved in comparison.
+The results of the experiment have demonstrated significant advantages of the new algorithm over other methods.
 
@@
 These functions can be interpreted as goals of decision making and, for certainty, we will assume that the less the function value, the better the result of achieving this goal.
+In other words, the best result for the $s$-th criterion corresponds to the global minimum of the function $\varphi_s(y)$ in the domain (\ref{eq:01}).
 
-Partial criteria form a vector-function called \textit{vector criterion}
+Partial criteria form a vector-function called \textit{vector criterion}. 
 \begin{equation}
@@
 
-If one or more partial criteria are not positive, this is not too restrictive, because it is possible to convert this problem to the statement satisfying conditions (\ref{eq:07}) with preserving the Pareto solution.
+If one or more partial criteria are not positive, this is not too restrictive, because it is possible to convert this problem to the statement satisfying conditions (\ref{eq:07}) with preserving the Pareto solution \cite{Pardalos2017}.
 

### 2023-10-17 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
fix two comments

@@
 
-If one or more partial criteria are not positive, this is not too restrictive, because it is possible to convert this problem to the statement satisfying conditions (\ref{eq:07}) with preserving the Pareto solution \cite{Pardalos2017}.
+If one or more partial criteria are not positive, this is not too restrictive, because it is possible to convert this problem to the statement satisfying conditions (\ref{eq:07}) with preserving the Pareto solution.
+A substantiation for such a transformation is given, for example, in \cite{Pardalos2017}.
 
@@
 
+\textbf{Remark}.
+The above method is based on the ISA  global search algorithm, which has an explicit geometric interpretation (see \cite{Strongin2000}).
+Using trial points, it is possible to design a minorant for the objective function $f(x)$.
+Then the value of characteristic, taken with the opposite sign, i.e.
+$-R(i)$, is the minimum value of the minorant in the interval $(x_{i-1}, x_i)$.
+Moreover, the point of a new trial $x^{p+1}$, computed according to (\ref{eq:26}), is carried out at the point, at which the minimum of the minorant is situated.
+
 \section{Experimental efficiency testing}\label{sec:4}

### 2023-10-17 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
R -> B f -> phi

@@
 The above method is based on the ISA  global search algorithm, which has an explicit geometric interpretation (see \cite{Strongin2000}).
-Using trial points, it is possible to design a minorant for the objective function $f(x)$.
+Using trial points, it is possible to design a minorant for the objective function $\varphi(x)$.
 Then the value of characteristic, taken with the opposite sign, i.e.
-$-R(i)$, is the minimum value of the minorant in the interval $(x_{i-1}, x_i)$.
+$-B(i)$, is the minimum value of the minorant in the interval $(x_{i-1}, x_i)$.
 Moreover, the point of a new trial $x^{p+1}$, computed according to (\ref{eq:26}), is carried out at the point, at which the minimum of the minorant is situated.

### 2023-10-18 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
phi -> f
fix remark

--- 2023-10-17 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
@@
 
-\textbf{Rule 4.} In the criteria space on the base of the marked up matrix $A^i_p$, a separating hyperplane $L_p$ is constructed applying logistic regression \cite{Yu2011} as efficient tool of machine learning.
+\textbf{Rule 4.} A hyperplane $\Pi_p$ is constructed that separates the point classes in the space of criteria labeled according to Rule 3.
+The classification algorithm based on logistic regression is used as a specific method of hyperplane construction \cite{Yu2011}.
+Note that these two classes are unbalanced in terms of the number of points, therefore, when building a hyperplane, non-dominated points are taken into account with the weight 0.95, and dominated ones with the weight 0.05.
 
-\textbf{Rule 5.} For each point $x_j, 1 \leq j \leq p$, the distance $d_j$ of the vector $(\varphi_1(x_j), \dots, \varphi_k(x_j))$ to the separating plane $L_p$ is determined.
+\textbf{Rule 5.} For each point $x_j, 1 \leq j \leq p$, the distance $d_j$ of the vector $(\varphi_1(x_j), \dots, \varphi_k(x_j))$ to the separating plane $\Pi_p$ is determined.
 This distance is considered to be negative if vector is placed above the plane and positive otherwise (see Fig.
@@
 \includegraphics[width=0.5\textwidth]{fig1}
-\caption{An example of estimating distances to the separating plane $L_p$.} \label{fig:1}
+\caption{An example of estimating distances to the separating plane $\Pi_p$.} \label{fig:1}
 \end{figure}
@@
 Then the value of characteristic, taken with the opposite sign, i.e.
-$-R(i)$, is the minimum value of the minorant in the interval $(x_{i-1}, x_i)$.
+$-B(i)$, is the minimum value of the minorant in the interval $(x_{i-1}, x_i)$.
 Moreover, the point of a new trial $x^{p+1}$, computed according to (\ref{eq:26}), is carried out at the point, at which the minimum of the minorant is situated.

### 2023-11-02 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
add letter sign

@@
 \usepackage{hyperref}
+
+\usepackage[misc,geometry]{ifsym}
 
@@
 %
-\author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
+\author{Konstantin Barkalov \Letter \orcidID{0000-0001-5273-2471} \and
 Vladimir Grishagin\orcidID{0000-0002-2884-3670} \and

### 2023-11-15 2023/MMST/samplepaper.tex
change grant number

@@
 \title{An algorithm for finding the global extremum\\ 
-of a partially defined function\thanks{This work was supported by the Russian Science Foundation, project No. 21-11-00204.}}
+of a partially defined function\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.
+FSWR-2023-0034, and by the Research and Education Mathematical Center ``Mathematics for Future Technologies''.}}
 %

### 2023-12-16 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
15 pages

@@
 
-The second approach applies meta-heuristic methods (genetic algorithms, simulation annealing, particle swarm optimization, artificial bee or ant colony, firefly algorithm, etc.) inspired by modeling the behavior of biological agents \cite{Nebro2009,RC05,Mostaghim2007,Durillo2010} or physical processes \cite{DPA02,ZLT01}.
+The second approach applies meta-heuristic methods (genetic algorithms, simulation annealing, particle swarm optimization, artificial bee or ant colony, firefly algorithm, etc.) inspired by modeling the behavior of biological agents \cite{NDG09,RC05,Mostaghim2007,Durillo2010} or physical processes \cite{DPA02,ZLT01}.
 Methods of this class are simple enough for implementation, but because of heuristic nature they do not guarantee convergence to global optimum in multiextremal case.
@@
 Vector
-\begin{equation}
-    \label{eq:06a}
-    \mu = (\mu_1, \dots, \mu_k) 
-\end{equation}
+$\mu = (\mu_1, \dots, \mu_k)$
 under conditions (\ref{eq:06}) belongs to a $(k-1)$-dimensional convex polyhedron in the Euclidean space  $R^k$.
@@
 
-Let the set $M$ contain the finite number $m$ of different weight vectors (\ref{eq:06a}):
+Let the set $M$ contain the finite number $m$ of different weight vectors:
 \begin{equation}
@@
 \end{equation}
-is computed, where
-\begin{equation}
-    \label{eq:21}
-\Gamma=
-\max_{1 \leq j \leq p-1} {\frac{|u_j - u_{j-1}|}{\rho_j} },
-\end{equation}
-\begin{equation}
-    \label{eq:22}
-\rho_j = \sqrt[n]{x_j-x_{j-1}},
-\end{equation}
+is computed, where $\Gamma = \max_{1 \leq j \leq p-1} {\frac{|u_j - u_{j-1}|}{\rho_j} }$, $\rho_j = \sqrt[n]{x_j-x_{j-1}}$,
 and $r>1$ is a parameter of the method.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{fig1}
+\includegraphics[width=0.6\textwidth]{fig1}
 \caption{An example of estimating distances to the separating plane $\Pi_p$.} \label{fig:1}
@@
 \centering
-\includegraphics[width=0.7\textwidth]{fig3}
+\includegraphics[width=0.6\textwidth]{fig3}
 \caption{Influence of the parameter $\beta$.} \label{fig:3}

### 2024-03-04 2023/ITMO paper/elsarticle-template-num.tex
fix s
submitted
some minor updates

@@
 %% Русский язык - закомментировать в финальной версии
-%\usepackage[russian]{babel}
+\usepackage[russian]{babel}
 
@@
 \usepackage{hyperref}
+\usepackage[normalem]{ulem}
 
@@
 The reduction is based on the use of space-filling curves (Peano curves).
-These approaches are implemented in the iOpt open-source framework of intelligent optimization methods.
-To demonstrate the advantages of iOpt, we compare it with the well-known Optuna and HyperOpt frameworks when tuning hyperparameters of various machine learning methods on a representative set of datasets.
+These approaches are implemented in the open-source framework of \textit{i}ntelligent \textit{opt}imization methods \textcolor[rgb]{1,0,0}{(iOpt)}.
+To demonstrate the advantages of iOpt, we compare it with the well-known Optuna and HyperOpt frameworks when tuning hyperparameters of various machine learning methods on \textcolor[rgb]{1,0,0}{\sout{a} the}  representative \textcolor[rgb]{1,0,0}{ \sout{set of} } datasets.
 The results show that Lipschitz global optimization methods provide comparable (in terms of quality) results in a significantly shorter time compared to known hyperparameter tuning algorithms.
@@
 %%Research highlights
-\begin{highlights}
-\item A framework of Lipschitz optimization methods is developed for the tuning of hyperparameters
-\item Dimensionality reduction based on Peano space-filling curves is used to solve multidimensional problems
-\item A scheme is proposed for reducing the problems, some of whose parameters are categorical, to a single problem with continuous parameters
-\item The developed methods (unlike the known methods of hyperparameter tuning) are deterministic, which ensures stable results at different runs of the algorithms
-
-\end{highlights}
+%\begin{highlights}
+%\item A framework of Lipschitz optimization methods is developed for the tuning of hyperparameters
+%\item Dimensionality reduction based on Peano space-filling curves is used to solve multidimensional problems
+%\item A scheme is proposed for reducing the problems, some of whose parameters are categorical, to a single problem with continuous parameters
+%\item The developed methods (unlike the known methods of hyperparameter tuning) are deterministic, which ensures stable results at different runs of the algorithms 
+%
+%\end{highlights}
 
@@
 
-\linenumbers
+%\linenumbers
 
@@
 
-In this paper, we present iOpt\footnote{\url{https://github.com/aimclub/iOpt/}}, a framework of intelligent optimization methods focused on solving problems of selecting optimal values of hyperparameters of various algorithms (including machine learning and heuristic optimization methods).
+In this paper, we present a framework of \textit{i}ntelligent \textit{opt}imization methods  \textcolor[rgb]{1,0,0}{iOpt\footnote{\url{https://github.com/aimclub/iOpt/}}}, focused on solving problems of selecting optimal values of hyperparameters of various algorithms (including machine learning \textcolor[rgb]{1,0,0}{and artificial intelligence} methods).
+
+\textcolor[rgb]{1,0,0}{Отличительной особенностью фреймворка iOpt является использование for hyperparameters tuning методов липшицевой оптимизации, которые являются детерминированными, обладают быстрой сходимостью, а также возможностью адаптации к различного рода специальным постановкам задач.
+Например, рассматриваемые методы допускают обобщение для работы с зашумленными функциями \cite{Sergeyev2020}.
+В данной работе они были адаптированы для решения HPO problems, в которых часть параметров часто являются дискретными и/или категориальными.}
+
 The paper is organized as follows.
@@
 It is known (see \cite{Strongin2000}) that if a multidimensional function  $\varphi(y), \; y \in D$,  satisfies the Lipschitz condition with the constant $L$, then the reduced one-dimensional function $\varphi(y(x)), \; x \in [0,1]$ satisfies the H{\"o}lder condition
-\[
+\begin{equation}\label{holder}
 \left|\varphi(y(x'))-\varphi(y(x''))\right|\leq H\left|x'-x''\right|^{1/N}, \; x',x''\in[0,1].
-\]
+\end{equation}
 Here, $N$ is the dimensionality of the original problem, and the coefficient
@@
 We have proposed and implemented a new deterministic approach to solving problems with mixed parameters; here we give a brief description of it. 
-Let the objective function of the problem depend on two parameter vectors: vector $y$, which belongs to the hyperinterval $D$, and vector $u$, which has a finite (and not very large) set of possible values, i.e.
+Let the objective function of the problem depend on two parameter vectors: vector $y$, which belongs to the hyperinterval $D$, and vector $u$, which has a finite (and not very large) set of values $U$, i.e.
 \begin{gather}\label{problem_i}
-\min{\left\{ \varphi(y,u) : y\in D \right\}},\\
+\min{\left\{ \varphi(y,u) : y\in D, u \in U \right\}},\\
 D=\left\{a_j \leq y_j \leq b_j, \; 1\leq j \leq N \right\}.\nonumber
@@
 
-Such finite sets of values can characterize many possible combinations of categorical hyperparameters of the machine learning algorithm under study.
+Such finite sets of values can characterize possible combinations of categorical hyperparameters of the machine learning algorithm under study.
 A typical example is the support vector machine, whose performance is influenced by the continuous regularization parameter $C$ and the kernel coefficient $\gamma$.
@@
 which translates any point of the interval $[0,S]$ to the region $D$ (the notation $E(x)$ corresponds to the integer part of the number $x$) and define the function
-\[
+\begin{equation}\label{f}
 f(x) = \varphi(Y(x),u_{E(x)+1}), x\in[0,S],
-\]
+\end{equation}
 which has, generally speaking, discontinuities at integer points $x_i = i, 1 \leq i \leq S-1$.
@@
 In this case, the categorical parameters can be assigned integer values, and the problem can be reduced to the formulation (\ref{problem_is1}).
-Since the solution of problems with partially integer parameters in accordance with the scheme described in Section \ref{sec_discr} implies the solution of problem (\ref{problem_is1}) with a discontinuous objective function, a modification of the algorithm that takes this feature into account is required.
+Since the solution of problems with partially integer parameters in accordance with the scheme described in Section \ref{sec_discr} implies the solution of problem (\ref{problem_is1}) with a discontinuous objective function, a modification of the global search algorithm that takes this feature into account is required.
 
 For this purpose, we introduce the classification of points from the search region using a special index $\nu_i=\nu(x_i)$.
-We will assume the indices of integer points $x_i = i, 0\leq i \leq S$, to be equal to zero, i.e.
-$\nu(x_i) = 0$.
+We will assume the indices of integer points $x_i = i, 0\leq i \leq S$, to be equal to zero.
 For all other values of $x\in(i,i+1),  0 \leq i < S$, we will assume the index value to be equal to unity, i.e.
-$\nu(x) = 1$.
-Then the result of the trial at some point $x^j\in[0,S]$  will be the set of values $(x^j, \nu^j=\nu(x^j), y^j=y(x^j), z^j = \varphi(y^j))$.
-In the case of  $\nu^j=0$, the value of $z^j$ will be considered undefined.
+\begin{equation}\label{index}
+	\nu(x) = \left\{
+   \begin{array}{lr}
+     0, \; x = i, 0\leq i \leq S, \\
+     1, \; x \neq i,  0 \leq i \leq S.
+   \end{array}
+\right.
+\end{equation}
+
+Then the result of the trial at some point $x^j\in[0,S]$  will be the set of values $(x^j, \nu^j=\nu(x^j), z^j = f(x^j))$. In the case of  $\nu^j=0$, the value of $z^j$ will be considered undefined.
 
@@
 1.
-At the initialization stage, the values $(x^i = i, \nu^i=0, y^i=y(x^i), z^i)$, $ i, 0\leq i \leq S$, i.e., the boundary points of the intervals $(x_{i-1},x_i), 1<i\leq S$, are entered into the set $\Omega_k$ containing the search  information.
+At the initialization stage, the values $(x^i = i, \nu^i=0, z^i)$, $ i, 0\leq i \leq S$, i.e., the boundary points of the intervals $(x_{i-1},x_i), 1<i\leq S$, are entered into the set $\Omega_k$ containing the search  information.
 After that, trials are carried out at arbitrary internal points (e.g., midpoints) of the intervals $(x_{i-1},x_i), 1<i\leq S$; the trial results are also added to the search information.
@@
 Here, $t$ is the number of the search interval that has the maximum characteristic.
-
 We note again that the characteristic $R(i)$ quantifies the possibility of finding a point $x^*$, which is the inverse image of the global optimizer $y^* = y(x^*)$, within the considered interval $(x_{i-1},x_i)$.
+
+Условия сходимости Global Search Algorithm для задач с Categorical Parameters (GSA-CP)  могут быть сформулированы в виде следующей теоремы.
+
+\textbf{Theorem} (convergence conditions)
+
+Пусть в постановке задачи (\ref{problem_is1}) функция $f(x)$ из (\ref{f}) удовлетворяет условию Гельдера с константой $H$ при $x \in (i, i + 1), 0 \leq i < S$.
+И пусть точка $\overline{x}$ be a limit point of the sequence $\left\{x^k\right\} $, полученной при минимизации функции $f(x)$ с помощью алгоритма GSA-CP. Then:
+\begin{enumerate}
+	\item 
+	если одноверменно с $\overline{x}$ существует другая предельная точка $x'$ последовательности $\left\{x^k\right\}$, тогда $f(\overline{x}) = f(x')$.
+	\item
+	$z^k=f(x^k) \geq f(\overline{x})$ при любом $k \geq 0$.
+	\item
+	Если на некотором шаге поиска величина $\mu$ from (\ref{mu}) satisfies the condition
+\begin{equation}\label{217}
+  r \mu > 4H,
+\end{equation}
+then $\overline{x}$ is a global minimizer of the function $f(x), x\in[0,S]$ and any global minimizer $x^\ast$ is also a limit point of the sequence $\left\{x^k\right\}$.
+\end{enumerate}
+
+\textbf{Proof}.
+Эта теорема может быть доказана аналогично теореме о сходимости index method of global optimization from \cite{Strongin2000}, принимая во внимание формулу (\ref{f}), определяющую функцию $f(x)$, правила (\ref{index}), задающие индексы у точек испытаний, а также взаимосвясь (\ref{holder}) между Lipschitz and H\"older conditions при редукции размерности с помощью evolvents.
 
@@
 \centering
-\caption{Time of hyperparameter tuning.}
+\caption{Time of hyperparameters tuning.}
 \label{tab:time}
@@
 \includegraphics[width=\textwidth]{fig_time}
-\caption{Time of hyperparameter tuning.} \label{fig8}
+\caption{Time of hyperparameters tuning.} \label{fig8}
 \end{figure}
@@
 The software that implements all of the described methods is available in the open repository \url{https://github.com/aimclub/iOpt/}.
-Python scripts that reproduce the results of the experiments described in this paper are available at \url{https://github.com/aimclub/iOpt/benchmarks}.
+Python scripts that reproduce the results of the experiments described in this paper are available in the benchmarks folder.
 

### 2024-03-07 2023/ITMO paper/elsarticle-template-num.tex
reformulate theorem in Russian

@@
 
-Let us number all the different combinations of categorical parameters with integers $s, 1\leq s \leq S$, i.e., let us assign a vector $u_s$ to each number $s$.
+Let us number all the different combinations of categorical parameters with integers $s = 1,...,S$, i.e., let us assign a vector $u_s$ to each number $s$.
 Then the problem under consideration can be written in the form
@@
 Using a dimensionality reduction scheme with Peano curves $y(x), x\in [0,1]$,  we can assign to each minimization problem on $y$ a one-dimensional minimization problem
-\[
- \min{\left\{ \varphi(y(x),u_s): x \in [0,1] \right\}}, s \in \{1,...,S\}.
-\]
+\begin{equation}\label{fi_s}
+ \min{\left\{ \varphi(y(x),u_s): x \in [0,1] \right\}}, s = 1,...,S.
+\end{equation}
 
@@
 
-Условия сходимости Global Search Algorithm для задач с Categorical Parameters (GSA-CP)  могут быть сформулированы в виде следующей теоремы.
-
-\textbf{Theorem} (convergence conditions)
-
-Пусть в постановке задачи (\ref{problem_is1}) функция $f(x)$ из (\ref{f}) удовлетворяет условию Гельдера с константой $H$ при $x \in (i, i + 1), 0 \leq i < S$.
-И пусть точка $\overline{x}$ be a limit point of the sequence $\left\{x^k\right\} $, полученной при минимизации функции $f(x)$ с помощью алгоритма GSA-CP. Then:
+%Новый текст на русском
+\textcolor[rgb]{1,0,0}{Условия сходимости Global Search Algorithm для задач с Categorical Parameters (GSA-CP) могут быть сформулированы в виде следующего утверждения.}
+
+\textbf{Утверждение} (сonvergence conditions)
+
+Пусть функции $ \varphi(y,u_s), y \in D, s = 1,...,S,$ form (\ref{problem_is}) удовлетворяют условию Липшица с соответствующими константами $L_s, s = 1,...,S$, и $L = \max\{L_s\}$.
+И пусть точка $\overline{x}$ be a limit point of the sequence $\left\{x^k\right\} $, полученной при минимизации функции $f(x)$ из (\ref{f}) с помощью алгоритма GSA-CP. Then:
 \begin{enumerate}
@@
 \begin{equation}\label{217}
-  r \mu > 4H,
+  r \mu > 2^{3-1/N} L \sqrt{N+3},
 \end{equation}
@@
 \textbf{Proof}.
-Эта теорема может быть доказана аналогично теореме о сходимости index method of global optimization from \cite{Strongin2000}, принимая во внимание формулу (\ref{f}), определяющую функцию $f(x)$, правила (\ref{index}), задающие индексы у точек испытаний, а также взаимосвясь (\ref{holder}) между Lipschitz and H\"older conditions при редукции размерности с помощью evolvents.
+Это утверждение доказывается аналогично теореме о сходимости index method of global optimization from \cite{Strongin2000}, принимая во внимание взаимосвясь (\ref{f}) функций  $f(x)$ и $\varphi(y,u_s)$, а также правила (\ref{index}), задающие индексы у точек испытаний.
 

### 2024-03-09 2024/NumMet/NMP_authorGuidelinesRu_2023.tex
first version of text in NumNet journal

% Windows: протестировано для MikTeX+TeXworks в режиме pdfLatex.
%
% Linux:
% Для получения pdf используйте команду  pdflatex rfa_2021.tex.
% Команда latex rfa_2021.tex выведет сообщение об ошибке.
%
\NeedsTeXFormat{LaTeX2e}
\documentclass[10pt,a4paper]{book}
\usepackage{NumMet_2023}
%--- Здесь можно вставить необходимые стили ---
% \usepackage{...}
%
%--- Здесь можно добавить свои команды --------
% \newcommand{}
\newcommand{\ceq}{\mathrel{\vcenter{\hbox{:=}}}}
%----------------------------------------------
\selectlanguage{russian}
\begin{document}
%===========================  ШАПКА СТАТЬИ =================================
\Article{Алгоритм липшицевой глобальной оптимизации\break
    для частично определенной функции}
    {Lipschitz global optimization algorithm \break
    of a partially defined function}

\Abstract{В статье обсуждается проблема нахождения глобального минимума липшецевой функции, которая может быть частично определена в области поиска. Такое явление возможно в силу особенностей оптимизируемого объекта и метода моделирования (например, численная нестабильность метода моделирования). В некоторых случаях эти подобласти известны, но в большинстве случаев информация о них отсутствует. В статье изложено краткое описание алгоритма глобального поиска для решения такого класса задач. Даны теоретические обоснования сходимости алгоритма. Проведены численные эксперименты, подтверждающие эффективность предложенного алгоритма.}
    {The paper discusses the problem of finding the global minimum of a Lipschitz function that may be partially defined in the search domain. This phenomenon is possible due to the nature of the optimized object and of the simulation method (for example, numerical instability of the simulation method). In some cases these subareas are known, but in most cases the information about them is missing. The paper provides a brief description of the global search algorithm for solving such class of problems. Theoretical justifications for the convergence of the algorithm are given. Numerical experiments confirming the efficiency of the proposed algorithm were carried out.}

\Keywords{липшецева глобальная оптимизация, 
        многоэкстремальные функции, 
        частично определенные функции.}
        {Lipschitz global optimization, 
        Multiextremal functions, 
        Partially defined functions.}

\Acknowledgements{Работа выполнена при поддержке Министерства науки и высшего образования Российской Федерации (проект \No~FSWR-2023-0034) и Научно-образовательного математического центра ``Математика для технологий будущего''.}
    {The work was supported by the Ministry of Science and Higher Education of the Russian Federation (project no.~FSWR-2023-0034) and by the Research and Education Mathematical Center ``Mathematics for Future Technologies''.}

\Citation{Усова~М.А., Баркалов~К.А.
    Алгоритм липшицевой глобальной оптимизации
    для частично определенной функции~//
    Вычислительные методы и программирование. 2024.
    \textbf{??}, \No~?. \pageref*{firstPage}--\pageref*{LastPage}.  
    doi 10.26089/NumMet.v??r???.}
    {M.~A.~Usova, K.~A.~Barkalov
    ``Lipschitz global optimization algorithm
    of a partially defined function'',
    Numerical Methods and Programming. \textbf{??} (?), 
    \pageref*{firstPage}--\pageref*{LastPage} (2024).
    doi~10.26089/NumMet.v??r???.}

\UDC{<указывает автор>}
\DOI{10.26089/NumMet.v??r???}% заполняет редакция
\VOL{??}{?}
\YEAR{2024}
\Received{16 июня 2024 г.}{June 16, 2024}
\Accepted{13 октября 2024 г.}{October 13, 2024}

\Author{М.~А.~Усова}{M.~A.~Usova}
\FullName{Марина~Андреевна~Усова}{Marina~A.~Usova}
\Institution{Национальный исследовательский Нижегородский государственный университет им.~Н.~И.~Лобачевского}
{National Research Lobachevsky State University of Nizhni Novgorod}
\Address{пр.Гагарина, 23}{Gagarina Prospekt, 23}
\Postcode{603022}
\City{Нижний Новгород}{Nizhni Novgorod}
\CountryOfResidence{Российская Федерация}{Russia}
%\Position{мл. науч. сотр.}{Assistant Research Fellow}
\Orcid{0000-0002-0722-6884}
\Email{usova@itmm.unn.ru}

\Author{К.~А.~Баркалов}{K.~A.~Barkalov}
\FullName{Константин~Александрович~Баркалов}{Konstantin~A.~Barkalov}
\Institution{Национальный исследовательский Нижегородский государственный университет им.~Н.~И.~Лобачевского}
{National Research Lobachevsky State University of Nizhni Novgorod}
\Address{пр.Гагарина, 23}{Gagarina Prospekt, 23}
\Postcode{603022}
\City{Нижний Новгород}{Nizhni Novgorod}
\CountryOfResidence{Российская Федерация}{Russia}
\AcademicDegree{д-р техн. наук.}{D.Eng.}
%\Position{мл. науч. сотр.}{Assistant Research Fellow}
\Orcid{0000-0001-5273-2471}
\Email{konstantin.barkalov@itmm.unn.ru}

\MakeArticleHeader
%========================= КОНЕЦ ШАПКИ СТАТЬИ ==============================

\pagebreak

\section{Введение}
Глобальная оптимизация занимается разработкой методов, призванных решать задачи отыскания точек абсолютного (глобального) минимума многоэкстремальных многомерных функций. Возможность достоверной оценки глобального оптимума в таких задачах принципиально основана на наличии априорной информации о решаемой задаче, позволяющей связать возможные значения целевой функции с известными значениями в точках проведенных поисковых испытаний.

Весьма часто такая информация представляется в виде предположения, что целевая функция $\phi(y)$ удовлетворяет условию Липшица с неизвестной априори константой $L$. Это предположение можно интерпретировать (применительно к прикладным задачам) как отражение ограниченности мощностей, порождающих изменения в моделируемой системе. При этом целевая функция зачастую задаётся в виде ``\textit{черного ящика}, является недифференцируемой и каждое вычисление ее значения в некоторой точке допустимой области может требовать значительных вычислительных ресурсов.

Для решения задач липшицевой глобальной оптимизации разработан ряд эффективных детерминированных методов 
[\ref{rfa:rulit:Grishagin2016_2},\ref{rfa:rulit:Jones2021},\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Birect2020},\ref{rfa:rulit:Sergeyev2017}]. Проведенные сравнения показывают, что детерминированные алгоритмы превосходят (по разным критериям) широко распространенные nature-inspired algorithms [\ref{rfa:rulit:Liberti2005},\ref{rfa:rulit:Sergeyev2018}].

Данная работа продолжает развитие одного из эффективных детерминированных методов решения задач липшицевой глобальной оптимизации – информационно-статистического алгоритма глобального поиска [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}]. Данный алгоритм предполагает решение многомерных задач посредством их сведения к задачам одномерной оптимизации при помощи space-filling curves (Peano curves).

Отметим, что для оптимизации сложных объектов реального мира естественным является использование при расчетах сложных математических моделей, что, как следствие, существенно увеличивает трудоемкость поиска оптимума. За последние десятилетия специалистами в области оптимизации и параллельных вычислений было предложено множество способов снижения вычислительной сложности и ускорения используемых алгоритмов, связанных как с решением возникающих оптимизационных задач [\ref{rfa:rulit:Kvasov2013},\ref{rfa:rulit:Sergeyev2020}], так и с численным анализом исходных моделей [\ref{rfa:rulit:Dongarra2022},\ref{rfa:rulit:Duwe2020}].

Однако в последнее время актуальной становится принципиально новая проблема прикладных задач – численная нестабильность исследуемых мо-делей в некоторых (заранее не известных) подобластях области изменения параметров. В таких подобластях невозможно корректно провести численное моделирование и вычислить значение целевой функции задачи. Данное явление можно интерпретировать либо как наличие в задаче некоторых скрытых ограничений (hidden constraints) [\ref{rfa:rulit:Stripinis2021}], или как наличие неизвестных разрывных областей [\ref{rfa:rulit:Audet2022}], или как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Sergeyev2003},\ref{rfa:rulit:Strongin2020}]. В такой постановке задача оптимизации существенно усложняется, т.к. область допустимых сочетаний параметров является заранее неопределенной.

В статье отражены результаты нового направления исследований, связан-ного с использованием импутации\footnote{Термин импутиция (imputation) в статье используется в соответствии с устоявшейся терминологией в области машинного обучения, связанной с восстановлением недостающих данных.} (восстановления) значений целевой функции в ходе работы алгоритма глобального поиска при решении задач с частично определенной целевой функцией. Предложенный подход основан на адаптации вычислительных правил алгоритма глобального поиска. Эффективность предложенного подхода демонстрируется численно.

\section{Постановка задачи}
В общем виде задача глобальной оптимизации может быть сформулирова-на следующим образом:
\begin{equation}\label{eq1} 
\phi^*=\phi(y^* )=\min_{y \in D} \phi(y), D=\left\{ y \in R^N: a_i \leq y_i \leq b_i, \; 1 \leq i \leq N \right\},
\end{equation}
где $y=(y_1,y_2,...,y_N)$ -– вектор варьируемых параметров, $D$ -- $N$-мерный гиперкуб, $N$ –- размерность решаемой задачи.
Используя кривые типа развертки Пеано, однозначно отображающие отрезок $[0,1]$ на $N$-мерный единичный гиперкуб
$$
D=\left\{ y \in R^N: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\},
$$
исходную задачу (\ref{eq1}) можно редуцировать к одномерной задаче
\begin{equation}\label{eq2} 
f^*(x)=\phi(y(x^* ))=\min_{x \in [0,1]} \left\{ \phi(y(x)) \right\},
\end{equation}
что позволяет применить для ее решения эффективные алгоритмы одномер-ной оптимизации.

О целевой функции задачи мы делаем следующие предположения.

Во-первых, целевая функция может быть многоэкстремальной, недиффе-ренцируемой и, более того, заданной в форме ``черного ящика'' (т.е. в виде некоторой подпрограммы, на вход которой подается аргумент, а выходом является соответствующее значение функции).

Во-вторых, каждое вычисление функции в некоторой точке допустимой области может требовать значительных вычислительных ресурсов.

В-третьих, целевая функция удовлетворяет условию Липщица
\begin{equation}\label{eq3} 
| \phi (y')-\phi (y'') | \leq L \| y'-y'' \|, \; y',y'' \in D,
\end{equation}
где $L$ -- константа Липщица. Известно, что схема редукции размерности с использованием кривых Пеано сопоставляет многомерной задаче с липшицевой целевой функцией (\ref{eq1}) задачу (\ref{eq2}) с одномерной целевой функцией, удовлетворяющей условию Гёльдера
\begin{equation}\label{eq4} 
| f(x')-f(x'') | \leq K \rho(x',x''), \; x',x'' \in [0,1],
\end{equation}
где $\rho(x',x'') =  |x' - x''|^{1/N}$ и $N$ -- размерность исходной многомерной задачи, а коэффициент $K$ связан с константой Липшица $L$ соотношением $K \leq 2L\sqrt {N+3}$ [\ref{rfa:rulit:Strongin2000}].

Различные варианты алгоритмов для решения такого класса задач и соот-ветствующая теория сходимости представлены в работах [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}].

В-четвертых, в некоторых подобластях области поиска $D$ (в частном случае, в одной или нескольких ее точках) целевая функция может быть не определена.

Исходя из опыта решения прикладных оптимизационных задач [\ref{rfa:rulit:Barkalov2022},\ref{rfa:rulit:Gubaydullin2022}], мы будем предполагать, что суммарный объем областей невычислимости составляет небольшую долю объема области поиска $D$ (около 10-20\%)\footnote{При тестировании алгоритма мы рассматривали и более сложные задачи, в которых объём невычислимой области достигал 50-80\%.Задачи такого рода также успешно решались предложенным алгоритмом, но для этого требовалось большое количество испытаний, поскольку метод пытался исследовать всю область невычислимости. В связи с этим в правила построения тестовых задач для размерности $N=5$ были внесены изменения таким образом, чтобы области невычислимости были не повсеместно по контуру, а только в угловых частях области поиска}.

Последнее предположение делает невозможными применение классического информационно-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска  [\ref{rfa:rulit:Strongin2000}] или других методов липшицевой оптимизации [\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Sergeyev2017}] и требует разработки модификации на случай решения задач с не всюду определенной целевой функцией.

\section{Метод решения задач с частично определенной целевой функцией}

\subsection{Алгоритм глобального поиска}

%В рамках предлагаемого подхода для решения задач (\ref{eq2}) используется информационно-статистическая теория глобального поиска. Основная идея алгоритма в использовании накопленной информации для определения очередного интервала, на котором абсолютный минимум наиболее вероятен. Очередная точка в этом интервале соответствует математическому ожиданию положения минимума.
Приведем более подробное описание \textbf{алгоритма глобального поиска} (АГП) для решения задач (\ref{eq2}).

На каждой итерации глобального поиска выполняется \textit{испытание}. Испытанием будем называть вычисление значения оптимизируемой функции $\phi (y(x))$ из (\ref{eq2}). Первое испытание будет осуществляться в серединной внутренней точке $x^1 \in (0,1)$. Выбор точки $x^{k+1}, k \geq 1$ очередного $(k+1)^\text{-го}$ испытания осуществляется на основе следующих правил.

\textit{Правило 1.} Перенумеровать (нижним индексом) точки $x^i, 0 \leq i \leq k$ предшествующих испытаний в порядке возрастания значений координаты, то есть
\begin{equation}\label{eq5} 
0=x_0 < x_1 < ... < x_i < ... < x_{k}=1
\end{equation}
и сопоставить им значения $z_i=f(x_i), 0 < i < k$, вычисленные в этих точках, и индекс $v_i=v(x_i)$, определяемый по правилу
\begin{equation}\label{eq6} 
v_i=v(x_i)=
  \begin{cases}
    -1 & {\quad \text{if } x_i \text{ -- граничная точка,}}\\
    1  & {\quad \text{if } x_i \text{ -- внутренняя точка.}}
  \end{cases}
\end{equation}
Точки $x_0=0$ и $x_{k}=1$ введены дополнительно (значения $z_0$ и $z_{k}$ не определены) для удобства последующих обозначений.

\textit{Правило 2.} Вычислить текущую нижнюю оценку
\begin{equation}\label{eq7} 
\mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i}, 2 \leq i \leq k \right\} , \Delta _i= (x_i-x_{i-1})^{1/N},
\end{equation}
для константы Гёльдера $K$ из (\ref{eq4}) для редуцированной функции $f(x)$, где $r>1$ параметр надёжности ; если $\mu=0$, то принять $\mu=1$.
\textit{Правило 3.} Определить текущее лучшее значение целевой функции
\begin{equation}\label{eq8} 
z^*=\min \left\{ f(x_i): 1\leq i \leq k \right\}.
\end{equation}

\textit{Правило 4.} Для каждого интервала $(x_{i-1},x_i),1 \leq i \leq k$, вычислить значение $R(i)$, называемое \textit{характеристикой} интервала, согласно выражению
\begin{equation}\label{eq9} 
R(x_i)=
  \begin{cases}
    \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1},\\
    2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1},\\
    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
  \end{cases}
\end{equation}

\textit{Правило 5.} Выбрать интервал $(x_{t-1},x_t)$ с максимальным значением характеристики $R(i)$:
\[
R(t)= \max\{R(i): \; 1 \leq i \leq k\}.
\]

\textit{Правило 6.} Провести очередное испытание в серединной точке интервала $(x_{t-1},x_t)$, если индексы его концевых точек не совпадают, т.е.
\begin{equation}\label{eq10} 
x^{k+1}=\frac {x_t+x_{t-1}}{2}, v(x_{t-1})\neq v(x_t).
\end{equation}
В противном случае провести испытание в точке 
\begin{equation}\label{eq11} 
x^{k+1}= \frac {x_t+x_{t-1}}{2} -  \text{sign} {(z_t-z_{t-1})} \frac{1}{2r} \left[\frac {{|z_t-z_{t-1}|}}{\mu} \right]^N, v(x_{t-1})=v(x_t).
\end{equation}

\textit{Условие остановки.} Поиск завершен, если длина интервала $(x_{t-1},x_t)$, содержащего очередную точку $x^{k+1}$, не превышает заданной точности $\varepsilon$, то есть $\Delta _t \leq \varepsilon$, где $t$ из Правила 5, а $\varepsilon>0$ -- задаваемый параметр алгоритма.

\subsection{Подходы к восстановлению значений целевой функции}
Описанная ранее проблема не всюду определенной целевой функции является частичным аналогом уже давно существующей проблемы пропущенных значений в Data Science, причинами которой могут быть, например, технические проблемы или сборка датасета из нескольких источников с различными наборами параметров.

Стоит отметить, хотя некоторые алгоритмы машинного обучения умеют принимать во внимание и даже восстанавливать пропущенные значения в данных, например, LightGBM имеет режим игнорирования пропущенных значений, а XGBoost восстанавливает данные за счёт уменьшения функции потерь при обучении, большинство моделей не могут обрабатывать пропущенные значения и требуют валидные данные без $NaN$ или «missing» значений, что делает данную проблему достаточно серьёзной.

Подходы к импутации (восстановлению) пропущенных значений в Data Science, предложенные специалистами в данной сфере, самые разные:

\begin{itemize}[itemsep=0pt,parsep=2pt,topsep=2pt,partopsep=0pt]
\item{удаление каждого наблюдения, содержащего одно или несколько пропущенных значений (крайнее средство, алгоритм теряет доступ к полезной информации, содержащейся в непропущенных значениях наблюдений);}
\item{замена пропущенных значений средним/медианой (не работает с качественными переменными, не учитывает корреляцию данных);}
\item{замена самым часто встречающимся значением или константой (отличается от использования медианы только работой с качественными параметрами);}
\item{замена данных, используя метод $k$ ближайших соседей или kNN (в отличие от первых учитывает корреляцию между параметрами, чувствителен к выбросам, вычислительно дорого);}
\item{множественная импутация данных (MICE) – импутация каждого значения проводится не один раз, а много (позволяет понять, насколько надёжно или ненадёжно предложенное значение);}
\item{импутация данных с помощью глубокого обучения (восстанавливает недостающие значения за счет тренировки нейронной сети на тех точках, для которых есть все параметры, на больших данных вычислительно дорого).}
\end{itemize}
Общим недостатком является отсутствие универсального алгоритма, необходимость для каждой конкретной задачи искать наиболее подходящие методы.

Для импутации значений целевой функции может рассматриваться в целом схожий набор подходов, однако учитывающий липшицевость функции и использующий информационно-статистическую теорию, что позволяет подобрать универсальный алгоритм для данного класса задач. Далее точки, в которых обнаружена невычислимость, интерпретируются как имеющие нулевой индекс и (\ref{eq6}) заменяется расширенной формулой

\begin{equation}\label{eq12} 
v_i=v(x_i)=
  \begin{cases}
    -1, & {\quad \text{if } x_i \text{ -- граничная точка}},\\
    0, & {\quad \text{if } x_i \text{ -- невычислимая точка}},\\
    1, & {\quad \text{if } x_i \text{ -- внутренняя точка}}.
  \end{cases}
\end{equation}

Опишем возможные подходы к реализации модификации.

\textbf{Первый подход.} Игнорировать отсутствие значения целевой функции в точке, расчёт характеристики интервала хотя бы с одной невычислимой точкой производить по длине интервала, т.е.
\begin{equation}\label{eq13} 
R(i)=\Delta _i, \text{ if } v(x_i)=0 \text{ или } v(x_{i-1})=0.
\end{equation}

\begin{figure}
\includegraphics[width=\textwidth]{pic/fig1.png}
\caption{Примеры распределения поисковых испытаний в областях невычислимости при первом подходе к импутации значений} \label{fig1}
\end{figure}

Существенным минусом данного подхода является большое число избыточных испытаний после попадания в область невычислимости за счёт большого значения характеристики, получаемой по формуле (\ref{eq13}).

\textbf{Второй подход.} Игнорировать отсутствие значения целевой функции в точке, расчёт характеристики интервала с двумя невычислимыми или с невычислимой и граничной точками производить по длине интервала, умноженной на понижающую константу, т.е.
\begin{equation}\label{eq14} 
R(i)={(1-{1}/{r})}^N \Delta _i,
\end{equation}
если $v(x_i)=v(x_{i-1})=0$ или $v(x_i)=0$ и $v(x_{i-1})=-1$ или $v(x_i)=-1$ и $v(x_{i-1})=0$.

Расчёт характеристики интервала с невычислимой и вычислимой точками производить по правилам работы с граничным интервалом, т.е.
\begin{equation}\label{eq15} 
R(i)=2\Delta _i-4 \frac {(z_i-z^*)}{r \mu},\text{ if } v(x_{i-1})=-1  \text{ or } v(x_{i-1})=0, v(x_i)=1,
\end{equation}
\begin{equation}\label{eq16} 
R(i)=2\Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu},\text{ if } v(x_{i-1})=1, v(x_i)=-1 \text{ or } v(x_i)=0.
\end{equation}

\begin{figure}
\includegraphics[width=\textwidth]{pic/fig2.png}
\caption{Примеры распределения поисковых испытаний в областях невычислимости при втором подходе к импутации значений} \label{fig2}
\end{figure}

Данный подход частично решает выявленные проблемы первого подхода, однако при усложнении тестовых задач не решает проблему избыточных испытаний и требует значительного увеличения параметра надёжности $r$. Примеры для одномерных задач с использованием первых двух подходов представлены на рис.~\ref{fig1} и рис.~\ref{fig2}, где черным показаны испытания, попавшие в область с неопределенными значениями, а зелёным вычислимые точки области поиска.

\textbf{Третий подход.} Восстанавливать значения целевой функции в точках для интервалов с двумя невычислимыми точками как среднее между соседними точками, т.е. и использовать соответствующее вычислительное правило из формулы (\ref{eq9}), считая точки вычислимыми. Если соседние точки также являются невычислимыми, использовать формулу (\ref{eq14}). В случае интервала с одной невычислимой точкой производить по формулам (\ref{eq15}) и (\ref{eq16}).

\begin{figure}
\centering
\includegraphics[width=\textwidth]{pic/fig3.png}
\caption{Импутация невычислимых значений средним соседних точек. Подход \No~3} \label{fig3}
\end{figure}

Существенными минусами будет не использование информации о координатах, а также редкость работы добавленного вычислительного правила за счёт поиска значений на замену только среди непосредственных соседей.

\textbf{Четвёртый подход.} Восстанавливать значения целевой функции в точках для интервалов с двумя невычислимыми точками как среднее между ближайшими вычислимыми точками с учётом положения координат, т.е.
\begin{equation}\label{eq17} 
%\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \rho(x,x')}{\rho(x',x'') },
\end{equation}
где $z'$ и $z''$ -- значения целевой функции в ближайших к интервалу $(x_{i-1},x_i)$ вычислимых точек $x'$ и $x''$, а $\rho(x_1,x_2) =  |x_1 - x_2|^{1/N}$. В остальном подход идентичен третьему. Формула (\ref{eq14}) в данном случае будет применяться лишь тогда, когда невычислимая область находится по соседству с граничной точкой.
\begin{figure}
\centering
\includegraphics[width=\textwidth]{pic/fig4.png}
\caption{Импутация невычислимых значений средним соседних точек. Подход \No~4} \label{fig4}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{pic/fig5.png}
\caption{Импутация невычислимых значений средним между минорантой и мажорантой. Подход \No~5} \label{fig5}
\end{figure}

\textbf{Пятый подход.} Чтобы максимально использовать поисковую информацию и предположения о виде функции, можно восстанавливать значения целевой функции в точках с учётом условия Гёльдера.

\textit{Шаг 1.} Найти крайние точки пересечения миноранты и мажоранты
\begin{equation}\label{eq18} 
\hat{x}',\hat{x}''=\frac {x'+x''}{2}\pm \text{sign}(z''-z')\cdot \frac {1}{2r} \cdot {\left(\frac {z''-z'}{\mu}\right)}^N,
\end{equation}

\textit{Шаг 2.} Восстановить значения целевой функции по правилу
\begin{equation}\label{eq19} 
\tilde{z}(x)=
  \begin{cases}
    z', & {\quad x \leq \hat{x}',}\\
    %\frac {z'+z''}{2}- \frac {r \mu}{2} (\| x - x' \| + \| x - x'' \|), & {\quad \hat{x}' < x < \hat{x}'',}\\
    \frac {z'+z''}{2}- \frac {r \mu}{2} (\rho(x,x')  + \rho(x,x'')), & {\quad \hat{x}' < x < \hat{x}'',}\\
    z'',  & {\quad x \geq \hat{x}'',}
  \end{cases}
\end{equation}
где $\rho(x_1,x_2) =  |x_1 - x_2|^{1/N}$.
Значение $\tilde{z}(x)$ может быть интерпретировано как среднее значение между минорантой и мажорантой для функции, удовлетворяющей условию Гёльдера.

Приведём описание модификации алгоритма глобального поиска на случай не всюду вычислимой функции на основе данного подхода.

\subsection{Описание модифицированного алгоритма глобального поиска}
Алгоритм остаётся прежним с внесением в него следующих небольших дополнений для вычислительных формул на случай попадания в невычислимые области поиска:

\begin{enumerate}[itemsep=0pt,parsep=2pt,topsep=2pt,partopsep=0pt]
  \item Индекс $v_i=v(x_i)$ определяется по формуле (\ref{eq12}).
  \item При вычислении текущей нижней оценки для константы Гельдера $K$ в формуле (\ref{eq7}) использовать только интервалы с двумя вычислимыми точками.
  \item Текущее лучшее значение целевой функции $z_v^*$ определяется для разных значений индекса
\begin{equation}\label{eq20} 
z_v^*=
  \begin{cases}
    -\varepsilon _r, & {\quad v=0 ,}\\
    \min \left\{ f(x_i): v(x_i)=1, 0 < i < k \right\}, & {\quad v=1 ,}
  \end{cases}
\end{equation}
где $\varepsilon _r > 0$ -- параметр метода.

  \item Для каждого интервала $(x_{i-1}, x_i), 1 \leq i \leq k$, the вычисление характеристики $R(i)$ происходит согласно следующим вычислительным формулам.

        \textit{Случай 1.} Для интервалов с \textbf{двумя вычислимыми точками} или \textbf{вычислимой и граничной точками} применить классические правила алгоритма глобального поиска
\begin{equation}\label{eq21} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z_1^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1,}\\
    2 \Delta _i-4 \frac {(z_i-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1,}\\
    2 \Delta _i-4 \frac {(z_{i-1}-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \textit{Случай 2.} Для интервалов с \textbf{граничной и невычислимой точками} произвести импутацию значения для невычислимой точки на основе соседней вычислимой точки и применить формулу
\begin{equation}\label{eq22} 
R(i)=
  \begin{cases}
    2 \Delta _i-4 \frac {(z_{i+1}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=0,}\\
    2 \Delta _i-4 \frac {(z_{i-2}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \textit{Случай 3.} Для интервалов с \textbf{невычислимой и вычислимой точками} произвести импутацию значения для невычислимой точки на основе соседней вычислимой точки и применить формулу
\begin{equation}\label{eq23} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-2})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-2}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=1,}\\
     \Delta _i+\frac {{(z_{i+1}-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_{i+1}+z_{i-1}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0.}
  \end{cases}
\end{equation}
        \textit{Случай 4.} Для интервалов с \textbf{двумя невычислимыми точками} произвести импутацию значений для невычислимой точки на основе правила (\ref{eq19}) и применить формулу
\begin{equation}\label{eq24} 
R(i)=\Delta _i+\frac {{(\tilde{z}_i-\tilde{z}_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {\tilde{z}_i+\tilde{z}_{i-1}-2z_1^*}{r \mu}, \; v(x_{i-1})=v(x_i)=0.
\end{equation}
        \textit{Случай 5.} В случае невозможности проведения любой импутации использовать формулу
\begin{equation}\label{eq25} 
R(i)=\Delta _i \cdot {\left( 1-\frac {1}{r} \right)}^N+z_0^*.
\end{equation}
  \item Если максимальную характеристику имеют несколько интервалов, отдавать предпочтение вычислимым интервалам с минимальным номером.
  \item Если лучшим интервалом оказался интервал с двумя невычислимыми точками, очередное испытание проводить в серединной точке интервала.
\end{enumerate}

\begin{nmAlgorithm}{pril:prog:alg1}
      {Функция $g$ (вычисление координат точки $v$ по ее порядковому номеру $k$)}
      {Function $g$ (calculating the coordinates of the point $v$ by its ordinal $k$)}
     \algLine{\algFUNC $g(k,d,p)$}
     \algLine[1]{$u_{n-1} \ceq \left\lfloor k/\left(d-1\right)^{n-2}\right\rfloor$}
     \algLine[1]{$u_n \ceq u_{n-1}$}
     \algLine[1]{$k\ceq k\mod (d-1)^{n-2}$}
     \algLine[1]{\algFOR $j=\left(n-3\right)\ldots0$ \algDO}
     \algLine[2]{$u_j\ceq \left\lfloor k/(d-1)^j\right\rfloor+1$}
     \algLine[2]{$k\ceq k\mod(d-1)^j$}
     \algLine[1]{\algENDFOR}
     \algLine[1]{\textbf{return} $(v_1,\dots,v_n)$}
     \algLine{\algENDFUNC}
\end{nmAlgorithm}

\subsection{Условия сходимости}

Установим условия сходимости для модификации алгоритма и покажем, что расширение алгоитма вычисления характеристики не рождает предельных точек другого типа по сравнению с предельными точками классического АГП.

\textbf{Теорема}. Формулировка теоремы.

\textbf{Доказательство}. Доказательство теоремы.

\section{Результаты вычислительных экспериментов}

Вычислительные эксперименты проводились на машине с процессором Intel\textregistered\ Core\texttrademark\ i7-10750H CPU @ 2.60GHz с использованием разработанного программного комплекса, реализующего описанную модификацию из Раздела 3. Для генерации серии задач использовался GKLS-генератор, описанный в [\ref{rfa:rulit:Gaviano2003}]. Он позволяюет порождать задачи многоэкстремальной оптимизации с заранее известными свойствами. Данные задачи были дополнены областями невычислимости разных видов.

Каждая серия экспериментов состоит из 100 задач GKLS со всюду вычислимой целевой функцией (original), 100 задач с невычислимостью на границе области поиска (round) и 100 задач со случайно заданными областями невычислимости (random).

\begin{figure}[h]
\includegraphics[width=\textwidth]{pic/fig6.png}
\caption{Примеры тестовых задач с \No~19, 25, и 35 из первой серии экспериментов} \label{fig6}
\end{figure}

В первой серии экспериментов проверялась корректность работы модификации и выявлялись особенности проведения испытаний в областях невычислимости. В эксперименте использовались задачи размерности $N=2$, параметр надежности был задан $r=5.5$, а точность $\varepsilon=0.01$. Результаты экспериментов представлены в табл.~\ref{tab1}.

На рис.~\ref{fig6} приведены примеры работы модификации с несколькими задачами серии. На рисунках синими точками отмечены испытания в вычислимых областях, красным найденный глобальный минимум задачи, а черным -- точки испытаний, проводившихся в области невычислимости. Также для наглядности области невычислимости окрашены в серый цвет.

Так как модификация с классической серией задач GKLS работает согласно базовому алгоритму глобального поиска, можно провести ее сравнение с классическим АГП. Так, для модификации сохранено общее поведение АГП и наблюдаются те же области сгущения испытаний вблизи локальных минимумов, что объясняется срабатыванием в большинстве случаев именно классических вычислительных правил (\ref{eq21}), в табл.~\ref{tab1} видно, что число испытаний, проводимых в невычислимых областях не велико. Общее число испытаний для классических задач и задач с невычислимостями соизмеримо, метод не проводит избыточных итераций в неперспективных областях благодаря расчётным формулам, использующим информацию о ближайших вычислимых точках.

\begin{table}[h]
\caption{Сравнение модификации алгоритма с классической версией алгоритма глобального поиска}\label{tab1}
\centering
\begin{tabular}{|l|c|c|c|}
\hline
 &  { original }  & { round }  & { random }  \\
\hline
Среднее число испытаний & 781.7	& 854.9 & 839.0 \\
\begin{tabular}{@{}l@{}} {Среднее число испытаний} \\  { в областях невычислимости } \end{tabular}  & - & 185.8 & 42.3 \\
{Число решенных задач} & 100 / 100 & 100 / 100  &  100 / 100 \\
\hline
\end{tabular}
\end{table}

Использование правила (\ref{eq25}) формирует сетку испытаний внутри граничных областей невычислимости, что хорошо заметно на задаче \No~35. При этом у границы невычислимой области близкой к потенциальному минимума формируется сгущение точек испытаний при попытке подойти к нему с разных сторон благодаря вычислительным правилам (\ref{eq22})--(\ref{eq24}), что наблюдается во всех приведенных в качестве примеров задачах.

Решаемость задач 100\%, однако для серии с граничной областью потребо-валось дорешивание одной задачи с немного большим значением $r=6.0$, для серии со случайными областями то же значение $r$ использовалось для решения двух задач.

\begin{table}
\caption{Серия задач разной размерности}\label{tab2}
\centering
\begin{tabular}{|l|l|c|c|c|}
\hline
 \begin{tabular}{@{}c@{}} {N} \\ {}\end{tabular} &
 \begin{tabular}{@{}c@{}} {Тип} \\ {тестовых задач} \end{tabular} &
 \begin{tabular}{@{}c@{}} {Среднее число испытаний} \\ {} \end{tabular} &
 \begin{tabular}{@{}c@{}} {Среднее число испытаний} \\ {в области невычислимости} \end{tabular} &
 \begin{tabular}{@{}c@{}} {Число решенных задач} \\ {}\end{tabular} \\
\hline
$3$ & original & 12251.1 & - & 100 / 100 \\
$ $ & round & 12201.8 & 4763.6 & 100 / 100 \\
$ $ & random & 12210.7 & 63.4 & 100 / 100 \\
\hline
$4$ & original & 25144.5 & - & 100 / 100 \\
$ $ & round & 29840.3 & 15004.4  & 100 / 100 \\
$ $ & random & 25287.5 & 26.4 & 100 / 100 \\
\hline
$5$ & original & 56725.9 & - & 100 / 100 \\
$ $ & round & 47840.1 & 24109.2 & 100 / 100\\
$ $ & random & 57874.8 & 82.6 & 100 / 100\\
\hline
\end{tabular}
\end{table}

Во второй серии экспериментов проводилось решение серий задач разных размерностей. В эксперименте использовались задачи размерности $N=3,4,5$, параметр надежности был задан $r=4.5$, а точность $\varepsilon=0.02$. Глобальный минимум считался найденным, если алгоритм генерировал точку испытания в $\varepsilon$-окрестности глобального минимума. Результаты выполненных экспериментов представлены в табл.~\ref{tab2}.

Как показывают результаты выполненных экспериментов, метод отлично справляется со случайными попаданиями в области невычислимости, разбросанными по области поиска случайно, а в случае больших граничных областей пытается провести полноценное исследование, не проводя при этом слишком много испытаний.

\section{Заключение}

В данной работе была рассмотрена задача нахождения глобального минимума дорогостоящей липшицевой функции вида ``черный ящик''. По сравнению с традиционной формулировкой, используемой в глобальной оптимизации, задача рассмотренная здесь имеет важное отличие: целевая функция может быть не определена в некоторых подобластях области поиска (в частном случае, в одной или нескольких ее точках). В таких подобластях невозможно корректно провести численное моделирование и вычислить значение целевой функции. Появление данной постановки обусловлено частой проблемой численной нестабильности исследуемых моделей прикладных задач.

Основная трудность заключается в невозможности в большинстве случаев заранее узнать о подобластях невычислимости, фактически, область допустимых сочетаний параметров является заранее неопределенной. Данное явление можно интерпретировать либо как наличие в задаче некоторых скрытых ограничений.

Исследуемая в данной работе проблема является частичным аналогом уже давно существующей в Data Science проблемы пропущенных значений. В качестве решения специалистами в данной сфере были предложены различные подходы к импутации (восстановлению) пропущенных значений. %Эти идеи нашли своё отражение при разработке нового метода.

Было проведено исследование применимости уже существующих подходов, дополненных различными вариантами учета липшицевости функции и использующих информационно-статистическую теорию. В результате удалось разработать универсальный эффективный алгоритм для данного класса задач. Построенный алгоритм является развитием одного из эффективных детерминированных методов решения задач липшицевой глобальной оптимизации – информационно-статистического алгоритма глобального поиска. В рамках модификации была произведена адаптация вычислительных правил алгоритма с внедрением алгоритма импутации значений.

Были проведены численные эксперименты на сериях многоэкстремальных задач GKLS размерностей $N = 2,3,4,5$. Результаты продемонстрировали надежность и эффективность предложенного подхода. Модификация отлично справляется со случайными попаданиями в подобласти невычислимости, разбросанными по области поиска. В случае больших граничных областей невычислимости метод всё также эффективно решает задачу, проводя полноценное исследование, в том числе, в невычислимых областях.

%=======================  Список литературы ==================================
%
% Команда \LITERRUS печатает "СПИСОК ЛИТЕРАТУРЫ" и обнуляет счетчики
%
\LITERRUS
% \rlitem{arg1}{arg2} --- создает пункт списка литературы.
% arg1 --- символическая ссылка на источник, используется при создании ссылки командой \ref
% arg2 --- текст, помещаемый в пункт списка. Может содержать команды выбора начертания 
%          символов и др.

%7
\rlitem{rfa:rulit:Grishagin2016_2}{%
\textit{Grishagin~V.A., Israfilov~R.A.}
Global search acceleration in the nested optimization scheme~// 
AIP Conference Proceedings. 2016. \textbf{1738}. 400010.
\doi{10.1063/1.4952198}.
}
%9
\rlitem{rfa:rulit:Jones2021}{%
\textit{Jones~D., Martins~J.}
The direct algorithm: 25 years later~// 
J. Glob. Optim. 2021. \textbf{79}, N~3. 521--566.
\doi{10.1007/s10898-020-00952-6}.
}
%12
\rlitem{rfa:rulit:PaulaviciusZilinskas2014}{%
\textit{Paulavi{\v c}ius~R. and {\v Z}ilinskas~J.}
Simplicial Global Optimization.
New York: Springer, 2014.
\doi{10.1007/978-1-4614-9093-7}.
}
%13
\rlitem{rfa:rulit:Birect2020}{%
\textit{Paulavi{\v c}ius~R., Sergeyev~Y.D., Kvasov~D.E., {\v Z}ilinskas~J.}
Globally-biased {BIRECT} algorithm with local accelerators for expensive global optimization~// 
Expert Syst. Appl. 2020. \textbf{144}. 113052.
\doi{10.1016/j.eswa.2019.113052}.
}
%14
\rlitem{rfa:rulit:Sergeyev2017}{%
\textit{Sergeyev~Y.D., Kvasov~D.E.}
Deterministic global optimization: An introduction to the diagonal approach.
New York: Springer, 2017.
\doi{10.1007/978-1-4939-7199-2}.
}
%11
\rlitem{rfa:rulit:Liberti2005}{%
\textit{Liberti~L., Kucherenko~S.}
Comparison of deterministic and stochastic approaches to global optimization~// 
Int. Trans. Oper. Res. 2005. \textbf{12}. 263--285.
}
%15
\rlitem{rfa:rulit:Sergeyev2018}{%
\textit{Sergeyev~Y.D., Kvasov~D.E., Mukhametzhanov~M.S.}
On the efficiency of nature-inspired metaheuristics in expensive global optimization with limited budget~// 
Sci. Rep. 2018. \textbf{8}, N~1. 435.
}
%16
\rlitem{rfa:rulit:Sergeyev2013}{%
\textit{Sergeyev~Y.D., Strongin~R.G., Lera~D.}
Introduction to Global Optimization Exploiting Space-Filling Curves.
New York: Springer Briefs in Optimization, 2013.
\doi{10.1007/978-1-4614-8042-6}.
}
%20
\rlitem{rfa:rulit:Strongin2000}{%
\textit{Strongin~R.G., Sergeyev~Y.D.}
Global optimization with non-convex constraints. Sequential and parallel algorithms.
Dordrecht: Kluwer Academic Publishers, 2000.
}
%10
\rlitem{rfa:rulit:Kvasov2013}{%
\textit{Kvasov~D.E., Sergeyev~Y.D.}
Lipschitz global optimization methods in control problems~// 
Autom. Remote Control. 2013. \textbf{74}, N~9. 1435--1448.
\doi{10.1134/s0005117913090014}.
}
%17
\rlitem{rfa:rulit:Sergeyev2020}{%
\textit{Sergeyev~Y.D., Candelieri~A., Kvasov~D.E., Perego~R.}
Safe global optimization of expensive noisy black-box functions in the $\delta$-Lipschitz framework~// 
Soft Comput. 2020. \textbf{24}, N~23. 17715--17735.
\doi{10.1007/s00500-020-05030-3}.
}
%4
\rlitem{rfa:rulit:Dongarra2022}{%
\textit{Dongarra~J.J.}
The evolution of mathematical software~// 
Commun. ACM. 2022. \textbf{65}, N~12. 66--72.
\doi{10.1145/3554977}.
}
%5
\rlitem{rfa:rulit:Duwe2020}{%
\textit{Duwe~K., et al.}
State of the art and future trends in data reduction for high-performance computing~// 
Supercomput. Front. Innov. 2020. \textbf{7}, N~1.
\doi{10.14529/jsfi200101}.
}
%19
\rlitem{rfa:rulit:Stripinis2021}{%
\textit{Stripinis~L., Paulavi{\v c}ius~R.}
A new {DIRECT}-{GLh} algorithm for global optimization with hidden constraints~// 
Optim. Lett. 2021. \textbf{15}, N~6. 1865--1884.
\doi{10.1007/s11590-021-01726-z}.
}
%1
\rlitem{rfa:rulit:Audet2022}{%
\textit{Audet~C., Batailly~A., Kojtych~S.}
Escaping unknown discontinuous regions in blackbox optimization~// 
SIAM J. Optim. 2022. \textbf{32}, N~3. 1843--1870.
\doi{10.1137/21m1420915}.
}
%3
\rlitem{rfa:rulit:Candelieri2019}{%
\textit{Candelieri~A.}
Sequential model based optimization of partially defined functions
under unknown constraints~// 
J. Glob. Optim. 2019. \textbf{79}, N~2. 281--303.
\doi{10.1007/s10898-019-00860-4}.
}
%18
\rlitem{rfa:rulit:Sergeyev2003}{%
\textit{Sergeyev~Y.D., Pugliese~P., Famularo~D.}
Index information algorithm with local tuning for solving multidimensional global optimization problems with multiextremal constraints~// 
Math. Program. 2003. \textbf{96}, N~23. 489--512.
\doi{10.1007/s10107-003-0372-z}.
}
%21
\rlitem{rfa:rulit:Strongin2020}{%
\textit{Strongin~R.G., Barkalov~K.A., Bevzuk~S.A.}
Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints~// 
Soft Comput. 2020. \textbf{24}, N~16. 11853--11865.
\doi{10.1007/s00500-020-05078-1}.
}
%2
\rlitem{rfa:rulit:Barkalov2022}{%
\textit{Barkalov~K.A., et al.}
On solving the problem of finding kinetic parameters of catalytic isomerization of the pentane-hexane fraction using a parallel global search algorithm~// 
Mathematics. 2022. \textbf{10}, N~19. 3665.
\doi{10.3390/math10193665}.
}
%8
\rlitem{rfa:rulit:Gubaydullin2022}{%
\textit{Gubaydullin~I.M., Enikeeva~L.V., Barkalov~K.A., Lebedev~I.G., Silenko~D.G.}
Kinetic modeling of isobutane alkylation with mixed c4 olefins and sulfuric acid as a catalyst using the asynchronous global optimization algorithm~// 
Commun. Comput. Inf. Sci. 2022. \textbf{1618}. 293--306.
\doi{10.1007/978-3-031-11623-0_20}.
}
%6
\rlitem{rfa:rulit:Gaviano2003}{%
\textit{Gaviano~M., Kvasov~D.E., Lera~D., Sergeyev~Y.D.}
Software for generation of classes of test functions with known local and global minima for global optimization~// 
ACM Trans. Math. Softw. 2003. \textbf{29}, N~4. 469--480.
}

%=======================  Список литературы ==================================
\medskip
\DateRU

\medskip

\MakeAuthorsInfoRU

%===========================  REFERENCES =====================================
\pagebreak

\REFERENCES

%7
\elitem{rfa:enlit:Grishagin2016_2}{%
V.~A.~Grishagin, R.~A.~Israfilov, ``Global search acceleration in the nested optimization scheme'',
AIP Conference Proceedings. {\bf 1738}, 400010 (2016). 
\doi{10.1063/1.4952198}.
}
%9
\elitem{rfa:enlit:Jones2021}{%
D.~Jones, J.~Martins, ``The direct algorithm: 25 years later'',
J. Glob. Optim. {\bf 79} (3), 521--566 (2021). 
\doi{10.1007/s10898-020-00952-6}.
}
%12
\elitem{rfa:enlit:PaulaviciusZilinskas2014}{%
R.~Paulavi{\v c}ius and J.~{\v Z}ilinskas
{\it Simplicial Global Optimization}
(Springer, New York, 2014).
\doi{10.1007/978-1-4614-9093-7}.
}
%13
\elitem{rfa:enlit:Birect2020}{%
R.~Paulavi{\v c}ius, Y.~D.~Sergeyev, D.~E.~Kvasov, J.~{\v Z}ilinskas, ``Globally-biased {BIRECT} algorithm with local accelerators for expensive global optimization'', 
Expert Syst. Appl. {\bf 144}, 113052 (2020).
\doi{10.1016/j.eswa.2019.113052}.
}
%14
\elitem{rfa:enlit:Sergeyev2017}{%
Y.~D.~Sergeyev, D.~E.~Kvasov
{\it Deterministic global optimization: An introduction to the diagonal approach}
(Springer, New York, 2017).
\doi{10.1007/978-1-4939-7199-2}.
}
%11
\elitem{rfa:enlit:Liberti2005}{%
L.~Liberti, S.~Kucherenko, ``Comparison of deterministic and stochastic approaches to global optimization'',
Int. Trans. Oper. Res. {\bf 12}, 263--285 (2005).
}
%15
\elitem{rfa:enlit:Sergeyev2018}{%
Y.~D.~Sergeyev, D.~E.~Kvasov, M.~S.~Mukhametzhanov, ``On the efficiency of nature-inspired metaheuristics in expensive global optimization with limited budget'',
Sci. Rep. {\bf 8} (1), 435 (2018).
}
%16
\elitem{rfa:enlit:Sergeyev2013}{%
Y.~D.~Sergeyev, R.~G.~Strongin, D.~Lera
{\it Introduction to Global Optimization Exploiting Space-Filling Curves}
(Springer Briefs in Optimization, Springer, New York, 2013).
\doi{10.1007/978-1-4614-8042-6}.
}
%20
\elitem{rfa:enlit:Strongin2000}{%
R.~G.~Strongin, Y.~D.~Sergeyev
{\it Global optimization with non-convex constraints. Sequential and parallel algorithms}
(Kluwer Academic Publishers, Dordrecht, 2000).
}
%10
\elitem{rfa:enlit:Kvasov2013}{%
D.~E.~Kvasov, Y.~D.~Sergeyev, ``Lipschitz global optimization methods in control problems'',
Autom. Remote Control. {\bf 74} (9), 1435--1448 (2013).
\doi{10.1134/s0005117913090014}.
}
%17
\elitem{rfa:enlit:Sergeyev2020}{%
Y.~D.~Sergeyev, A.~Candelieri, D.~E.~Kvasov, R.~Perego, ``Safe global optimization of expensive noisy black-box functions in the $\delta$-Lipschitz framework'',
Soft Comput. {\bf 24} (23), 17715--17735 (2020).
\doi{10.1007/s00500-020-05030-3}.
}
%4
\elitem{rfa:enlit:Dongarra2022}{%
J.~J.~Dongarra, ``The evolution of mathematical software'', 
Commun. ACM. {\bf 65} (12), 66--72 (2022).
\doi{10.1145/3554977}.
}
%5
\elitem{rfa:enlit:Duwe2020}{%
K.~Duwe, et al., ``State of the art and future trends in data reduction for high-performance computing'',
Supercomput. Front. Innov. {\bf 7} (1), (2020).
\doi{10.14529/jsfi200101}.
}
%19
\elitem{rfa:enlit:Stripinis2021}{%
L.~Stripinis, R.~Paulavi{\v c}ius, ``A new {DIRECT}-{GLh} algorithm for global optimization with hidden constraints'',
Optim. Lett. {\bf 15} (6), 1865--1884 (2021).
\doi{10.1007/s11590-021-01726-z}.
}
%1
\elitem{rfa:enlit:Audet2022}{%
Audet~C., Batailly~A., Kojtych~S., ``Escaping unknown discontinuous regions in blackbox optimization'',
SIAM J. Optim. {\bf 32} (3), 1843--1870 (2022).
\doi{10.1137/21m1420915}.
}
%3
\elitem{rfa:enlit:Candelieri2019}{%
A.~Candelieri, ``Sequential model based optimization of partially defined functions under unknown constraints'',
J. Glob. Optim. {\bf 79} (2), 281--303 (2019).
\doi{10.1007/s10898-019-00860-4}.
}
%18
\elitem{rfa:enlit:Sergeyev2003}{%
Y.~D.~Sergeyev, P.~Pugliese, D.~Famularo, ``Index information algorithm with local tuning for solving multidimensional global optimization problems with multiextremal constraints'', 
Math. Program. {\bf 96} (23), 489--512 (2003).
\doi{10.1007/s10107-003-0372-z}.
}
%21
\elitem{rfa:enlit:Strongin2020}{%
R.~G.~Strongin, K.~A.~Barkalov, S.~A.~Bevzuk, ``Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints'',
Soft Comput. {\bf 24} (16), 11853--11865 (2020).
\doi{10.1007/s00500-020-05078-1}.
}
%2
\elitem{rfa:enlit:Barkalov2022}{%
K.~A.~Barkalov, et al., ``On solving the problem of finding kinetic parameters of catalytic isomerization of the pentane-hexane fraction using a parallel global search algorithm'',
Mathematics. {\bf 10} (19), 3665 (2022).
\doi{10.3390/math10193665}.
}
%8
\elitem{rfa:enlit:Gubaydullin2022}{%
I.~M.~Gubaydullin, L.~V.~Enikeeva, K.~A.~Barkalov, I.~G.~Lebedev, D.~G.~Silenko, ``Kinetic modeling of isobutane alkylation with mixed c4 olefins and sulfuric acid as a catalyst using the asynchronous global optimization algorithm'',
Commun. Comput. Inf. Sci. {\bf 1618}, 293--306 (2022).
\doi{10.1007/978-3-031-11623-0_20}
}
%6
\elitem{rfa:enlit:Gaviano2003}{%
M.~Gaviano, D.~E.~Kvasov, D.~Lera, Y.~D.~Sergeyev, ``Software for generation of classes of test functions with known local and global minima for global optimization'',
ACM Trans. Math. Softw. {\bf 29} (4), 469--480 (2003).
}

\medskip
\DateEN

\bigskip

\MakeAuthorsInfoEN

%\end{document}
%============================ КОНЕЦ СТАТЬИ ==================================

%%%%%%%% В КОНЦЕ УДАЛИТЬ

%\section{Рекомендации к содержанию и структуре статей}

%Авторам оригинальных статей рекомендуется объем до 20 страниц.

%Объем аннотации не должен превышать 10--15 строк. Разбивать аннотации на абзацы не рекомендуется.

%Авторы статьи указывают УДК (для статей на русском языке), которым соответствует публикация.

%Рекомендуется для лучшего восприятия разбивать текст на абзацы размером не более 10--15~строк.

%Список литературы является неотъемлемой частью статьи и оформляется непосредственно в тексте статьи: использование отдельных файлов с расширениями bib, bbl и других не допускается.

%Авторы статей, публикуемых в журнале на русском языке, должны предоставить переведенные на английский язык:
%\begin{itemize}[itemsep=0pt,parsep=0pt,topsep=0pt,partopsep=0pt]
%\item фамилию, имя, отчество, ученую степень, ученое звание и должность каждого автора;
%\item название, почтовые адреса организаций, являющихся основными местами работы авторов, и их должности в этих организациях;
%\item название статьи;
%\item текст аннотации;
%\item ключевые слова;
%\item слова благодарности;
%\item подписи к рисункам, таблицам, листингам программ и алгоритмов, а также всю текстовую информацию, содержащуюся на рисунках, в таблицах и пр.
%\end{itemize}
%\vspace*{2mm}

%Между инициалами и фамилией всегда должен ставится пробел (А.А.~Иванов).
%В заголовке статьи пробелы также ставятся между инициалами: А.~А.~Иванов.

%Точка не ставится после заголовков разделов, названий таблиц и подписей рисунков, после ряда сокращенных наименований: с~--- секунда, г~--- грамм, мин~--- минута, сут~--- сутки, млн~--- миллион, млрд~--- миллиард и т.п. Точка ставится в конце последнего предложения аннотации, в конце списка ключевых слов, после сокращений: мес.~--- месяц, г.~--- год и др.

%Для набора выключных формул рекомендуется использовать окружение {\tt equation} с автоматической сквозной нумерацией. Для подавления нумерации используйте окружение {\tt equation*}.

%\pagebreak
%\smallskip

%\item Слова ``Теорема'', ``Следствие'', ``Лемма'', ``Утверждение'', ``Предложение'' помещаются в текст статьи явно и выделяются прямым полужирным шрифтом, а соответствующие формулировки печатаются прямым шрифтом.
%\item Слова ``Доказательство'', ``Определение'', ``Примечание'', ``Замечание'' выделяются прямым полужирным шрифтом, а соответствующие формулировки печатаются прямым шрифтом.
%\item Рекомендуется выполнять сквозную нумерацию арабскими цифрами для теорем, %лемм, определений и т.д.

%Допускается воспроизводить небольшие по объему (рекомендуется не более 1 страницы) фрагменты программ, иллюстрирующие важные для понимания статьи алгоритмы, средства языков программирования и т.п. Текстовые сообщения и комментарии в листингах должны быть представлены на английском языке.

%Рекомендуются следующие размеры иллюстраций: а)~шириной не более 165~мм и высотой не более 210~мм (в ширину колонки текста); б)~шириной не более 80 мм и высотой не более 140~мм (в половину ширины колонки); в) высотой 145~мм и шириной 240~мм (иллюстрация, развернутая на $90^\circ$ на всю страницу).
%Принимаются иллюстрации трех видов: штриховые рисунки (графики, диаграммы, графы и пр.), скриншоты экрана компьютера (например, цветовая визуализация поля, полученная в результате математического моделирования) и фотографии.
%Штриховые рисунки принимаются в виде файлов формата PNG (растр) разрешением 600 DPI при масштабе рисунка 1:1 или в формате EPS (векторные рисунки). Толщина самой тонкой линии на рисунке не должна быть меньше 5~px. Рекомендуется для основных линий использовать черный цвет. Важные для понимания элементы рисунка должны выделяться двумя-тремя цветами, которые контрастно воспроизводятся при черно-белой печати. Цвет фона рекомендуется белый. Фотографии принимаются в форматах JPEG, PNG, TIFF разрешением 300 DPI. Изображение на фотографии должно быть резким, контрастным. Цифровая обработка изображения не должна вызывать сомнений в достоверности фотографии. Надписи на рисунках должны быть выполнены на английском языке. Рекомендуется важные элементы рисунка обозначать цифрами и давать расшифровку в подписи. Серия иллюстраций предназначается для сравнения графиков, оценки влияния параметров, объяснения трендов, динамики и т.п. Рисунки, входящие в серию, должны быть одинакового размера и разрешения. Каждая иллюстрация в серии должна быть обозначена буквой английского алфавита.
%Редакция просит предоставлять наряду с файлами EPS исходные, необработанные файлы в форматах TIFF, PNG, JPEG.

%Подписи к рисункам, заголовки таблиц, листингов программ и алгоритмов должны быть выполнены на русском и английском языках. Для этого в стилевом файле определены команды \verb!\bicaption!.

%Список литературы формируется в порядке обращения к источникам в       тексте статьи.

%Для уменьшения вероятности ошибок в ссылках по тексту статьи необходимо пользоваться командами \verb!\label!, автоматически сохраняющей значения счетчиков для рисунков, таблиц, списков в символической переменной, и \verb!\ref!, которая создает ссылку по имени переменной.

%Необходимые документы
%Для публикации статьи авторы должны предоставить: 
% сопроводительное письмо авторов (Приложение~1 в этом документе) в электронном виде и его отсканированный вариант с ``живыми'' подписями;
% электронную \LaTeX--версию статьи (на русском или/и английском языках), подготовленную в соответствии с требованиями к оформлению статей, а также pdf--файл статьи и рисунки;
% текстовый файл, содержащий следующую информацию на русском и английском языках: а)~фамилия, имя, отчество, ученая степень, ученое звание, ORCID и e-mail каждого из авторов; б)~названия и почтовые адреса организаций, являющихся основными местами работы авторов и их должности в этих организациях; в)~название статьи; г)~текст аннотации; д)~ключевые слова.
%Автор взаимодействует с редакцией в основном через сайт журнала \url{https://num-meth.ru}. После заполнения регистрационной формы автор (один из авторов) должен:
%\item загрузить через форму все подготовленные необходимые документы и компоненты статьи в виде ZIP--архива, выбрав при загрузке {\tt Компонент статьи = ``ZIP-архив с исходными текстами (.tex и требуемые дополнительные файлы)''};
%\item дать описание архива, авторство и т.п. (рекомендуется сделать на этапе проверки деталей отправляемого файла);
%\item подтвердить отправку архива;
%\item ввести метаданные статьи (обязательно заголовок, аннотацию и ключевые слова) и информацию об авторе и всех соавторах (вся информация должна заноситься на русском и дублироваться на английском языке);
%\item подтвердить внесенные данные.

%%%%%%%% В КОНЦЕ УДАЛИТЬ

%============================ ПРИЛОЖЕНИЕ 6 ==================================
\newpage\nolinenumbers
%
% ПРИЛОЖЕНИЕ 4
% используйте как шаблон письма в редакцию
%
\thispagestyle{empty}
%
% Следующие две строки нужно убрать
%
%\appendix{Сопроводительное письмо в редакцию}{pril:sopr}
%

{\large
\hfill\parbox[t]{80mm}{
Главному редактору научного журнала\\[2mm]
``Вычислительные методы и\\
программирование''\\[2mm]
чл.-корр. РАН, проф. Вл. В. Воеводину
}
\vspace*{12mm}
\begin{center}
СОПРОВОДИТЕЛЬНОЕ ПИСЬМО
\end{center}

\vspace*{3mm}

Просим опубликовать в журнале ``Вычислительные методы и программирование'' статью 
``<НАЗВАНИЕ СТАТЬИ>''.

Сообщаем следующую информацию о каждом авторе статьи.

1. Фамилия, имя, отчество полностью (с написанием на английском языке).

2. Основное место работы полностью с указанием полного почтового адреса и URL сайта организации (с переводом на английский язык).

3. Должность по основному месту работы (с переводом на английский язык).

4. Ученая степень и ученое звание (с переводом на английский язык).

5. Телефон служебный, домашний и мобильный.

6. E-mail.

7. ORCID (\corr{обязательно}; если еще не получен, посетите \url{https://orcid.org/register}).

8. Номер и название рубрики ГРНТИ (только для статей на русском) и код OECD для подаваемой статьи.

9. Желательно: ссылки на профили авторов в elibrary.ru для упрощения и корректности подлинковки статьи.

\vspace*{2mm}
Текущую переписку по вопросам публикации статьи следует вести с <Фамилия И.О.>.

\vspace*{2mm}
Авторы согласны с правилами подготовки статьи к публикации, включая рецензирование, 
научное и литературное редактирование и доведение статьи до редакторских стандартов, 
принятых в рамках журнала.

Авторы ознакомлены с Публикационной этикой журнала и согласны с ее положениями.

Авторы также согласны с передачей журналу своего права на издание и распространение 
статьи в электронной и бумажной версиях, в том числе на размещение библиографической 
информации о статье в Российском индексе научного цитирования (РИНЦ) и в других базах 
научного цитирования и на размещение полных текстов статей в Научной электронной 
библиотеке (elibrary.ru) и портале http://www.mathnet.ru для свободного доступа всем пользователям Интернет независимо от их категории и местоположения.

\vspace*{15mm}
<Дата>\hfill\parbox[t]{70mm}{\hrulefill~<И.О. Фамилия>}

\vspace*{8mm}

\hfill\parbox[t]{70mm}{\hrulefill~<И.О. Фамилия>}

}%large

\end{document}

### 2024-03-11 2023/ITMO paper/elsarticle-template-num.tex
Описание экспериментов на суррогатных моделях

@@
 \subsection{Experimental setup}
+\label{sec_setup}
 
@@
 
+\subsection{Результаты вычислительных экспериментов на суррогатных моделях}
+
+Для проведения вычислительных экспериментов на основе суррогатных моделей нами использовался тот же набор алгоритмов оптимизации и параметры запуска, что и в случае реальных наборов данных (см.
+раздел \ref{sec_setup}).
+В качестве критерия оптимальности была выбрана метрика ``valid-acc''.
+В процессе оптимизации варьировались LearningRate, WeightDecay, а также функция активации в приделах, рекомендуемых авторами jahs\_bench [XXX].
+Для оптимизации алгоритмам максимальное число испытаний было ограничено 200.
+Для недетерминированных алгоритмов данные усреднялись по 10 запускам.
+Результаты вычислительных экспериментов представлены в таблице \ref{tab:met-surr}.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{graphicx}
+\begin{table}
+\centering
+\caption{Values of optimized metrics for the algorithms averaged over 10 runs (surrogate models)}
+\label{tab:met-surr}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{ccccccc}
+\hline
+\textbf{}                      & \textbf{iOpt}   & \textbf{Optuna(random)} & \textbf{Optuna(tpe)}     & \textbf{Optuna(cmaes)}   & \textbf{Optuna(nsgaii)} & \textbf{Hyperopt(tpe)} \\ \hline
+\textbf{cifar10}               & \textbf{88.394} & 88.144 $\pm$ 0.21\%         & 88.287 $\pm$ 0.21\%          & 88.218 $\pm$ 0.15\%          & 88.063 $\pm$ 0.27\%         & 88.316 $\pm$ 0.11\%        \\
+\textbf{fashion\_mnist}        & 94.797          & 94.677 $\pm$ 0.18\%         & 94.837 $\pm$ 0.13\%          & \textbf{94.952 $\pm$ 0.08\%} & 94.631 $\pm$ 0.35\%         & 94.759 $\pm$ 0.19\%        \\
+\textbf{colorectal\_histology} & \textbf{94.447} & 94.013 $\pm$ 0.33\%         & \textbf{94.447 $\pm$ 0.17\%} & 94.194 $\pm$ 1.18\%          & 94.150 $\pm$ 0.36\%         & 94.233 $\pm$ 0.17\%        \\ \hline
+\end{tabular}%
+}
+\end{table}
+Исходя из таблицы видно превосходство iOpt на моделях cifar10 и colorectal\_histology.
+При использовании моделей fashion\_mnist и фреймворк iOpt демонстрирует схожее качество.
+Отметим, что при использовании суррогатных моделей время вычисления критерия близко к нулю.
+Как следствия результаты, связанные с временем оптимизации, мы не приводим.
+
 \section{Conclusion}

### 2024-03-17 2024/ITMO paper/sn-article.tex
resubmission of ITMO paper in 2024

--- 2024-03-07 2023/ITMO paper/elsarticle-template-num.tex
@@
-%% 
-%% Copyright 2007-2020 Elsevier Ltd
-%% 
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%%
+%Version 3 October 2023
+% See section 11 of the User Manual for version history
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                                 %%
+%% Please do not use \input{...} to include other tex files.       %%
+%% Submit your LaTeX manuscript as one .tex document.              %%
+%%                                                                 %%
+%% All additional figures and files should be attached             %%
+%% separately and not embedded in the \TeX\ document itself.       %%
+%%                                                                 %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%% Template article for Elsevier's document class `elsarticle'
-%% with numbered style bibliographic references
-%% SP 2008/03/01
-%%
-%% 
-%%
-%% $Id: elsarticle-template-num.tex 190 2020-11-23 11:12:32Z rishi $
-%%
-%%
-\documentclass[preprint,12pt]{elsarticle}
+%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing
 
-%% Use the option review to obtain double line spacing
-%% \documentclass[preprint,review,12pt]{elsarticle}
+%%=======================================================%%
+%% to print line numbers in the margin use lineno option %%
+%%=======================================================%%
 
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn]{elsarticle}
-%% \documentclass[final,3p,times]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn]{elsarticle}
-%% \documentclass[final,5p,times]{elsarticle}
-%% \documentclass[final,5p,times,twocolumn]{elsarticle}
+%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
 
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
+%%======================================================%%
+%% to compile with pdflatex/xelatex use pdflatex option %%
+%%======================================================%%
 
-%% The amssymb package provides various useful mathematical symbols
-\usepackage{amssymb}
-%% The amsthm package provides extended theorem environments
-%%\usepackage{amsthm}
+%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
 
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
-\usepackage{lineno}
+%%Note: the following reference styles support Namedate and Numbered referencing.
+By default the style follows the most common style.
+To switch between the options you can add or remove Numbered in the optional parenthesis.
+%%The option is available for: sn-basic.bst, sn-vancouver.bst, sn-chicago.bst%
+
+%%\documentclass[sn-nature]{sn-jnl}% Style for submissions to Nature Portfolio journals
+%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+%\documentclass[sn-mathphys-num]{sn-jnl}% Math and Physical Sciences Numbered Reference Style 
+\documentclass[iicol]{sn-jnl}
+%%\documentclass[sn-mathphys-ay]{sn-jnl}% Math and Physical Sciences Author Year Reference Style
+%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
+%%\documentclass[sn-vancouver,Numbered]{sn-jnl}% Vancouver Reference Style
+%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style 
+%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
+
+%%%% Standard Packages
+
+%%<additional latex packages if required can be included here>
+
+\usepackage{graphicx}%
+\usepackage{multirow}%
+\usepackage{amsmath,amssymb,amsfonts}%
+\usepackage{amsthm}%
+\usepackage{mathrsfs}%
+\usepackage[title]{appendix}%
+\usepackage{xcolor}%
+\usepackage{textcomp}%
+\usepackage{manyfoot}%
+\usepackage{booktabs}%
+\usepackage{algorithm}%
+\usepackage{algorithmicx}%
+\usepackage{algpseudocode}%
+\usepackage{listings}%
 
@@
 \usepackage[russian]{babel}
-
-%% advanced mathematics
-\usepackage{amsmath}
-\usepackage[ruled]{algorithm2e}
-\usepackage{multirow}
-
-\usepackage{graphicx}
-\usepackage[table,xcdraw]{xcolor}
-\usepackage{amssymb}
-
-\usepackage{hyperref}
+%% Зачеркивание - закомментировать в финальной версии
 \usepackage[normalem]{ulem}
 
-\journal{Knowledge-Based Systems}
+\usepackage[round]{natbib}
+\bibliographystyle{abbrvnat}
+%\setcitestyle{authoryear,open={(},close={)}} %Citation-related commands
+
+%%%%
+
+%%%%%=============================================================================%%%%
+%%%%  Remarks: This template is provided to aid authors with the preparation
+%%%%  of original research articles intended for submission to journals published 
+%%%%  by Springer Nature. The guidance has been prepared in partnership with 
+%%%%  production teams to conform to Springer Nature technical requirements. 
+%%%%  Editorial and presentation requirements differ among journal portfolios and 
+%%%%  research disciplines. You may find sections in this template are irrelevant 
+%%%%  to your work and are empowered to omit any such section if allowed by the 
+%%%%  journal you intend to submit to. The submission guidelines and policies 
+%%%%  of the journal take precedence. A detailed User Manual is available in the 
+%%%%  template package for technical guidance.
+%%%%%=============================================================================%%%%
+
+%% as per the requirement new theorem styles can be included as shown below
+\theoremstyle{thmstyleone}%
+\newtheorem{theorem}{Theorem}%  meant for continuous numbers
+%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
+%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
+\newtheorem{proposition}[theorem]{Proposition}% 
+%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
+
+\theoremstyle{thmstyletwo}%
+\newtheorem{example}{Example}%
+\newtheorem{remark}{Remark}%
+
+\theoremstyle{thmstylethree}%
+\newtheorem{definition}{Definition}%
+
+\raggedbottom
+%%\unnumbered% uncomment this for unnumbered level heads
 
@@
 
-\begin{frontmatter}
-
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \affiliation{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%% \fntext[label3]{}
-
-\title{On hyperparameter tuning through Lipschitz global optimization} % and dimensionality reduction}?
+\title[Article Title]{On hyperparameter tuning through Lipschitz global optimization}
 %% О настройке гиперпараметров с помощью липшицевой глобальной оптимизации
 
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \affiliation[label1]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%%
-%% \affiliation[label2]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
+%%=============================================================%%
+%% GivenName	-> \fnm{Joergen W.}
+%% Particle	-> \spfx{van der} -> surname prefix
+%% FamilyName	-> \sur{Ploeg}
+%% Suffix	-> \sfx{IV}
+%% \author*[1,2]{\fnm{Joergen W.} \spfx{van der} \sur{Ploeg} 
+%%  \sfx{IV}}\email{iauthor@gmail.com}
+%%=============================================================%%
 
-\author[UNN,ITMO]{Konstantin Barkalov\corref{cor}}
-\cortext[cor]{Corresponding author: \url{konstantin.barkalov@itmm.unn.ru}}
-\author[UNN,ITMO]{Denis Karchkov}
-\author[UNN,ITMO]{Evgeny Kozinov}
-\author[UNN,ITMO]{Ilya Lebedev}
-\author[UNN,ITMO]{Denis Rodionov}
-\author[UNN,ITMO]{Marina Usova}
+\author*[1,2]{\fnm{Konstantin} \sur{Barkalov}}\email{konstantin.barkalov@itmm.unn.ru}
 
-\affiliation[UNN]{organization={Lobachevsky University},%Department and Organization
-            addressline={Gagarin Ave. 23}, 
-            city={Nizhni Novgorod},
-            postcode={603022}, 
-            %state={},
-            country={Russia}}
+\author[1,2]{\fnm{Denis} \sur{Karchkov}}
 
-\affiliation[ITMO]{organization={ITMO University},%Department and Organization
-            addressline={Lomonosova St. 9}, 
-            city={St. Petersburg},
-            postcode={191002}, 
-            %state={},
-            country={Russia}}
+\author[1,2]{\fnm{Evgeny} \sur{Kozinov}}
 
-\begin{abstract}
-%% Text of abstract
+\author[1,2]{\fnm{Ilya} \sur{Lebedev}}
 
-The quality of machine learning methods is substantially affected by their hyperparameters, while the evaluation of the quality criterion is a time-consuming operation.
+%\author[1,2]{\fnm{Denis} \sur{Rodionov}}
+
+\author[1,2]{\fnm{Marina} \sur{Usova}}
+
+\affil[1]{
+%\orgdiv{Department}, 
+\orgname{Lobachevsky University}, \orgaddress{\street{Gagarin Ave. 23}, \city{Nizhni Novgorod}, \postcode{603022},
+%\state{State},
+\country{Russia}}}
+
+\affil[2]{
+%\orgdiv{Department},
+\orgname{ITMO University}, \orgaddress{\street{Lomonosova St. 9}, \city{St. Petersburg}, \postcode{191002}, 
+%\state{State},
+\country{Russia}}}
+
+%%==================================%%
+%% Sample for unstructured abstract %%
+%%==================================%%
+
+\abstract{The quality of machine learning methods is substantially affected by their hyperparameters, while the evaluation of the quality criterion is a time-consuming operation.
 Therefore, it is important to develop intelligent methods for selecting optimal values of hyperparameters that require a small number of search trials.
@@
 To demonstrate the advantages of iOpt, we compare it with the well-known Optuna and HyperOpt frameworks when tuning hyperparameters of various machine learning methods on \textcolor[rgb]{1,0,0}{\sout{a} the}  representative \textcolor[rgb]{1,0,0}{ \sout{set of} } datasets.
-The results show that Lipschitz global optimization methods provide comparable (in terms of quality) results in a significantly shorter time compared to known hyperparameter tuning algorithms.
+The results show that Lipschitz global optimization methods provide comparable (in terms of quality) results in a significantly shorter time compared to known hyperparameter tuning algorithms.}
 
-\end{abstract}
+%%================================%%
+%% Sample for structured abstract %%
+%%================================%%
 
-%%Graphical abstract
-%\begin{graphicalabstract}
-%\includegraphics{grabs}
-%\end{graphicalabstract}
+% \abstract{\textbf{Purpose:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations.
+As a guide the abstract should not exceed 200 words.
+Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
+% 
+% \textbf{Methods:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations.
+As a guide the abstract should not exceed 200 words.
+Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
+% 
+% \textbf{Results:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations.
+As a guide the abstract should not exceed 200 words.
+Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
+% 
+% \textbf{Conclusion:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations.
+As a guide the abstract should not exceed 200 words.
+Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.}
 
-%%Research highlights
-%\begin{highlights}
-%\item A framework of Lipschitz optimization methods is developed for the tuning of hyperparameters
-%\item Dimensionality reduction based on Peano space-filling curves is used to solve multidimensional problems
-%\item A scheme is proposed for reducing the problems, some of whose parameters are categorical, to a single problem with continuous parameters
-%\item The developed methods (unlike the known methods of hyperparameter tuning) are deterministic, which ensures stable results at different runs of the algorithms 
-%
-%\end{highlights}
+\keywords{Lipschitz optimization, Global optimization, Multiextremal functions, Hyperparameter tuning}
 
-\begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
-Lipschitz optimization \sep Global optimization \sep Multiextremal functions \sep Hyperparameter tuning
+%%\pacs[JEL Classification]{D8, H51}
 
-%% PACS codes here, in the form: \PACS code \sep code
+%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}
 
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-
-\end{keyword}
-
-\end{frontmatter}
-
-%\linenumbers
-
-%% main text
-\section{Introduction}
-\label{sec_intro}
+\maketitle
+\section{Introduction}\label{sec_intro}
 
 Currently, complex scientific and/or industrial problems are most often solved by numerical methods.
-Most of such methods, for example, machine learning methods, have hyperparameters, and the choice of specific values of these hyperparameters can determine the efficiency (in a given metric) of the obtained solution under certain constraints on the computation time \cite{Hutter2019,nikitin2021}.
-The fact that in many cases the difference in the efficiency of the solution obtained with different values of hyperparameters can be quite significant, has resulted in the emergence of the hyperparameter optimization (HPO) problem, as well as the research of approaches to its solution that are implemented in the frameworks for automatic selection of hyperparameters \cite{Tune, optuna, hyperopt,Sherpa}.
+Most of such methods, for example, machine learning methods, have hyperparameters, and the choice of specific values of these hyperparameters can determine the efficiency (in a given metric) of the obtained solution under certain constraints on the computation time \citep{Hutter2019,nikitin2021}.
+The fact that in many cases the difference in the efficiency of the solution obtained with different values of hyperparameters can be quite significant, has resulted in the emergence of the hyperparameter optimization (HPO) problem, as well as the research of approaches to its solution that are implemented in the frameworks for automatic selection of hyperparameters \citep{Tune,optuna,hyperopt,Sherpa}.
 From the mathematical point of view, the HPO problem corresponds to a global optimization problem with fixed bounds on the variation of hyperparameters.
@@
 \textcolor[rgb]{1,0,0}{Отличительной особенностью фреймворка iOpt является использование for hyperparameters tuning методов липшицевой оптимизации, которые являются детерминированными, обладают быстрой сходимостью, а также возможностью адаптации к различного рода специальным постановкам задач.
-Например, рассматриваемые методы допускают обобщение для работы с зашумленными функциями \cite{Sergeyev2020}.
+Например, рассматриваемые методы допускают обобщение для работы с зашумленными функциями \citep{Sergeyev2020}.
 В данной работе они были адаптированы для решения HPO problems, в которых часть параметров часто являются дискретными и/или категориальными.}
@@
 
-\section{Related work}
-\label{sec_rel}
+\section{Related work}\label{sec_rel}
 
@@
 
-The simplest way to find the global optimum is the exhaustive search method (search on a uniform \cite{Bao2006} or random \cite{Bergstra2012} grid). 
-By selecting the number of analyzed variants for each of the hyperparameters, searching all the resulting combinations and choosing the best of them in a given metric, we obtain an estimate of the optimum with a certain accuracy \cite{Nevendra2022}.
+The simplest way to find the global optimum is the exhaustive search method (search on a uniform \citep{Bao2006} or random \citep{Bergstra2012} grid). 
+By selecting the number of analyzed variants for each of the hyperparameters, searching all the resulting combinations and choosing the best of them in a given metric, we obtain an estimate of the optimum with a certain accuracy \citep{Nevendra2022}.
 The exhaustive search method is easy to implement, it has perfect parallelism, but it has a significant drawback (an exponential growth of the number of grid nodes as the number of hyperparameters increases).
@@
 
-Metaheuristic (genetic and similar) algorithms are an alternative to the exhaustive search when the number of hyperparameters increases \cite{Opara2019}.
-These algorithms are widely used in the absence of a formula-based description of the model under study (``black-box'' model), which is typical for the considered HPO problems \cite{Zhou2021,Yang2022}.
+Metaheuristic (genetic and similar) algorithms are an alternative to the exhaustive search when the number of hyperparameters increases \citep{Opara2019}.
+These algorithms are widely used in the absence of a formula-based description of the model under study (``black-box'' model), which is typical for the considered HPO problems \citep{Zhou2021,Yang2022}.
 Metaheuristic algorithms only weakly depend on the number of hyperparameters, they use information from previous iterations to perform the current one, but due to the randomness inherent in the methods, they guarantee the finding of the global optimum only in the probabilistic sense.
 
-Another approach to finding the global optimum is Bayesian optimization, which is also used to solve ``black-box'' problems \cite{Frazier2018,Archetti2019}.
+Another approach to finding the global optimum is Bayesian optimization, which is also used to solve ``black-box'' problems \citep{Frazier2018,Archetti2019}.
 In fact, Bayesian optimization methods build a stochastic model of the function being optimized.
@@
 
-In well-known hyperparameter tuning frameworks, the Tree-Structured Parzen Estimator (TPE) algorithm is widely used \cite{hyperopt,NIPS2011}.
+In well-known hyperparameter tuning frameworks, the Tree-Structured Parzen Estimator (TPE) algorithm is widely used \citep{hyperopt,NIPS2011}.
 TPE is an iterative algorithm, conceptually close to Bayesian optimization methods.
@@
 
-The good results achieved by Bayesian optimization \cite{Joy2020} and TPE \cite{Watanabe2022a,Watanabe2022b} methods are based on the a priori assumption that the function corresponds to a certain stochastic model, such as a Gaussian process \cite{Rasmussen2005}.
+The good results achieved by Bayesian optimization \citep{Joy2020} and TPE \citep{Watanabe2022a,Watanabe2022b} methods are based on the a priori assumption that the function corresponds to a certain stochastic model, such as a Gaussian process \citep{Rasmussen2005}.
 However, there are other assumptions about the shape of the function in global optimization that yield remarkable results.
@@
 In this case, the function is said to satisfy the Lipschitz condition, and the problem to be solved is called a Lipschitz global optimization problem.
-For the class of Lipschitz global optimization problems, a number of efficient algorithms have been developed \cite{Jones2021,Paulavicius2020,Strongin2020,Sergeyev2017,PaulaviciusZilinskas2014}, which surpass many other global optimization methods \cite{Sergeyev2018} and have shown their applicability in a number of applications \cite{Kvasov2008,CANDELIERI2019,Gubaydullin2021}.
+For the class of Lipschitz global optimization problems, a number of efficient algorithms have been developed \citep{Jones2021,Paulavicius2020,Strongin2020,Sergeyev2017,PaulaviciusZilinskas2014}, which surpass many other global optimization methods \citep{Sergeyev2018} and have shown their applicability in a number of applications \citep{Kvasov2008,CANDELIERI2019,Gubaydullin2021}.
 
-The iOpt framework contains an implementation of Lipschitz optimization algorithms, which are an extension of the information-statistical global search algorithm described in detail in \cite{Strongin2000,Sergeyev2013}.
+The iOpt framework contains an implementation of Lipschitz optimization algorithms, which are an extension of the information-statistical global search algorithm described in detail in \citet{Strongin2000,Sergeyev2013}.
 In particular, the methods have been adapted to solve problems with partially integer parameters, which are very common in HPO problems.
-The experimental results presented in Section \ref{sec_exp} show that the iOpt framework is not inferior in HPO problems to well-known frameworks having the same purpose (such as Optuna \cite{optuna} and HyperOpt \cite{hyperopt}).
+The experimental results presented in Section \ref{sec_exp} show that the iOpt framework is not inferior in HPO problems to well-known frameworks having the same purpose (such as Optuna \citep{optuna} and HyperOpt \citep{hyperopt}).
 
-\section{Lipschitz optimization} 
-\label{sec_lip}
+\section{Lipschitz optimization}\label{sec_lip}
 
@@
 \begin{equation} \label{f_lip}
-	\left| \varphi(y')-\varphi(y'') \right| \leq L\left\| y' - y''  \right\| , \; y',y'' \in D, \; 0<L<\infty.
+	\left| \varphi(y')-\varphi(y'') \right| \leq L\left\| y' - y''  \right\| , \,
 \end{equation}
-Here $ \left\| \cdot \right\|$ denotes the Euclidean norm, and the constant $L$ is not known a priori and is subject to estimation in the process of solution.
+where $y',y'' \in D, \; 0<L<\infty$, and $ \left\| \cdot \right\|$ is denotes the Euclidean norm, and the constant $L$ is not known a priori and is subject to estimation in the process of solution.
 
@@
 Moreover, it is assumed that each search trial (i.e., computing the value of the function at a point in the admissible region) requires significant computational resources.
-This problem formulation fully corresponds to the task of tuning the hyperparameters of machine learning methods \cite{Joy2020,Wang2021}.
+This problem formulation fully corresponds to the task of tuning the hyperparameters of machine learning methods \citep{Joy2020,Wang2021}.
 
@@
 
-  By virtue of these inequalities, the values of the function at the points of the interval $[x', x'']$ must lie inside the region bounded by the straight lines passing through the points $(x', f(x'))$ and $(x'', f(x''))$ with slopes $L$ and $-L$ (see Fig.
-\ref{fig1}).
+  By virtue of these inequalities, the values of the function at the points of the interval $[x', x'']$ must lie inside the region bounded by the straight lines passing through the points $(x', f(x'))$ and $(x'', f(x''))$ with slopes $L$ and $-L$ (see Fig.~\ref{fig1}).
 
@@
 \centering
-\includegraphics[width=0.8\textwidth]{Fig1.pdf}
+\includegraphics[width=0.4\textwidth]{Fig1.pdf}
 \caption{The values of the function $f(x)$ must be inside the area highlighted by colour.} \label{fig1}
@@
 \[
-F(x) = \max\left\{f(x') - L(x-x'),f(x') + L(x-x'')\right\}, \; x\in [x', x''].
+F(x) = \max\left\{f(x') - L(x-x'),f(x') + L(x-x'')\right\},
 \] 
+where $\; x\in [x', x'']$.
 
@@
 Lipschitz global optimization methods use this idea in their computational rules; the proof of their convergence is also based on this idea (see, e.g.,
-\cite{Jones2021,PaulaviciusZilinskas2014,Sergeyev2013,Evtushenko2013}).
+\citet{Jones2021,PaulaviciusZilinskas2014,Sergeyev2013,Evtushenko2013}).
 
@@
 
-A well-known approach to solving the multidimensional problem (\ref{f_func}) is to reduce it to a one-dimensional problem and then apply efficient one-dimensional global optimization algorithms (see \cite{Strongin2000,Sergeyev2013}).
+A well-known approach to solving the multidimensional problem (\ref{f_func}) is to reduce it to a one-dimensional problem and then apply efficient one-dimensional global optimization algorithms (see \citet{Strongin2000,Sergeyev2013}).
 A well-proven approach here is to reduce the dimensionality of the problem by means of a Peano-Hilbert curve $y(x), x \in [0, 1]$.
@@
 
-It is known (see \cite{Strongin2000}) that if a multidimensional function  $\varphi(y), \; y \in D$,  satisfies the Lipschitz condition with the constant $L$, then the reduced one-dimensional function $\varphi(y(x)), \; x \in [0,1]$ satisfies the H{\"o}lder condition
+It is known (see \citet{Strongin2000}) that if a multidimensional function  $\varphi(y), \; y \in D$,  satisfies the Lipschitz condition with the constant $L$, then the reduced one-dimensional function $\varphi(y(x)), \; x \in [0,1]$ satisfies the H{\"o}lder condition
 \begin{equation}\label{holder}
-\left|\varphi(y(x'))-\varphi(y(x''))\right|\leq H\left|x'-x''\right|^{1/N}, \; x',x''\in[0,1].
+\left|\varphi(y(x'))-\varphi(y(x''))\right|\leq H\left|x'-x''\right|^{1/N}.
 \end{equation}
-Here, $N$ is the dimensionality of the original problem, and the coefficient
+Here, $x',x''\in[0,1]$, $N$ is the dimensionality of the original problem, and the coefficient
 $ H=2 L \sqrt{N+3}$.
 
-Remark 1.
+\begin{remark}
 The Lipshitz condition will not be satisfied for the reduced one-dimensional function $\varphi(y(x))$.
@@
 \end{figure}   
+\end{remark}
 
-Remark 2.
+\begin{remark}
 The theoretical Peano-Hilbert curve $y(x)$ is defined as a limiting object.
 Numerical algorithms use evolvents $y_m(x)$, that approximate the true curve with a given density level $2^{-m}$, which depends on the required search accuracy.
-Efficient schemes for computing several types of evolvents are described in detail in \cite{Sergeyev2013}.
+Efficient schemes for computing several types of evolvents are described in detail in \citet{Sergeyev2013}.
 For illustration, Fig.~\ref{fig:Peano} shows graphs corresponding to the evolvents $y_m(x)$ with density $m=4$ for dimensions $N=2$ and $N=3$.
+\end{remark}
 
-\subsection{The problem of discrete parameters}
-\label{sec_discr}
+\subsection{The problem of discrete parameters}\label{sec_discr}
 
 Of particular interest are problems where some of the parameters can assume values from a predetermined finite set, since it is more difficult to construct estimates of the optimum for these problems than for problems with continuous parameters.
-Many publications address the methods for solving mixed-parameter problems (see, for example, reviews \cite{Burer2012,Boukouvala2016}).
-Known deterministic methods for solving problems of this class are centered on solving linear or convex problems \cite{Lee2012}.
-It is suggested that complex multiextremal problems should be solved using heuristic approaches \cite{Belotti2013}.
+Many publications address the methods for solving mixed-parameter problems (see, for example, reviews \citet{Burer2012,Boukouvala2016}).
+Known deterministic methods for solving problems of this class are centered on solving linear or convex problems \citep{Lee2012}.
+It is suggested that complex multiextremal problems should be solved using heuristic approaches \citep{Belotti2013}.
 
@@
 %Целочисленные параметры = непрерывные параметры с округлением, добавить комментарий об этом?
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=1.0\textwidth]{scheme.pdf}
+\caption{iOpt general structure.} \label{fig_iOpt}
+\end{figure*}
 
@@
 
-\section{iOpt framework} 
-\label{sec_iOpt}
+\section{iOpt framework}\label{sec_iOpt}
 
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\textwidth]{scheme.pdf}
-\caption{iOpt general structure.} \label{fig_iOpt}
-\end{figure}
-
 The iOpt framework can be used to develop specialized decision support systems for solving the tasks of selecting optimal parameters for complex models or methods.
@@
 
-\subsection{Global search algorithm}
-\label{sec_GSA}
+\subsection{Global search algorithm}\label{sec_GSA}
 
@@
 It is convenient to represent the search information accumulated after conducting $k$ trials as a set in which the corresponding entries are renumbered (by the lower index) in ascending order of $x$, i.e., as a set
-\begin{equation}\label{omega}
-\Omega_k = \left\{  (x_i, y_i=y(x_i), z_i = \varphi(y_i)): 1 \leq i \leq k  \right\},	
-\end{equation}
-where $x_1 < x_2 < ... < x_k$.
+\begin{multline}\label{omega}
+\Omega_k = \left\{  (x_i, y_i=y(x_i), z_i = \varphi(y_i)): 1 \leq i \leq k  \right\},\\
+x_1 < x_2 < ... < x_k.
+\end{multline}
 
@@
 \end{equation}
-
 Here, $r>1$ is the algorithm parameter, $\Delta_i=(x_i-x_{i-1})^{1/N}$.
@@
 with the largest characteristic. If the right point of this interval has the number $t$ in the set $\Omega_k$, then the formula for calculating $x^{k+1}$ will look as follows:
-\begin{equation}\label{xk1}
-x^{k+1} = \frac{x_t+x_{t-1}}{2}- \frac{\mathrm{sign}(z_t-z_{t-1})}{2r} \left(\frac{\left|z_t-z_{t-1}\right|}{\mu}\right)^N.   
-\end{equation}
+\begin{multline}\label{xk1}
+x^{k+1}=\frac{x_t+x_{t-1}}{2}-\\
+-\frac{\mathrm{sign}(z_t-z_{t-1})}{2r}\left(\frac{\left|z_t-z_{t-1}\right|}{\mu}\right)^N.   
+\end{multline}
 
 Algorithm \ref{alg_GSA} shows the pseudo code of the Global Search Algorithm.
+
+\begin{algorithm}
+\caption{Global search algorithm}\label{alg_GSA}
+\begin{algorithmic}[1]
+\Require $\varphi(y), D, r, \epsilon, K_{max}$
+\Ensure $y_{min}$
+\State $k \Leftarrow 2$, $t \Leftarrow 2$
+\State $v_1 \Leftarrow (x_1 \Leftarrow 0, y_1 \Leftarrow y(x_1), z_1 \Leftarrow \varphi(y_1))$
+\State $v_2 \Leftarrow (x_2 \Leftarrow 1, y_2=y(x_2), z_2 \Leftarrow \varphi(y_2))$
+\State $\Omega_k \Leftarrow \left\{ v_1, v_2 \right\}$
+\While{ $\Delta_t \geq \epsilon$ and $k \leq K_{max}$}
+    \State compute the estimate of the H\"older constant $\mu$ by (\ref{mu})
+    \For{$i=0$ to $k$}
+        \State compute the characteristic $R(i)$ by (\ref{R})
+    \EndFor
+    \State determine the number $t$, for which $R(t) = \max \left\{ R(i), 1 \leq i \leq k \right\}$
+    \State compute the new trial point $x^{k+1}$ by (\ref{xk1})
+    \State perform a new trial at the point $x^{k+1}$
+    \State $\Omega_k \Leftarrow \Omega_k \cup (x^{k+1}, y^{k+1}, z^{k+1})$
+    \State $k \Leftarrow k + 1$
+\EndWhile
+\State $l \Leftarrow \arg \min \left\{ z_i, 1 < i \leq k \right\}$
+\State $y_{min} \Leftarrow y_l$
+\State \Return {$y_{min}$}
+\end{algorithmic}
+\end{algorithm}
+
 The input data of the algorithm are the objective function $\varphi(y)$ and the search region $D$ from (\ref{f_func}) and (\ref{f_D}), the algorithm parameter $r>1$, the minimum search accuracy $\epsilon > 0$, and the maximum allowable number of trials $K_{max}$.
@@
 
-\begin{algorithm}
-\LinesNumbered
- \KwIn{$\varphi(y), D, r, \epsilon, K_{max}$}
- \KwOut{$y_{min}$}
- Initialize $k=2$, $t=2$, $\Omega_k= \left\{ (x_1=0, y_1=y(x_1), z_1=\varphi(y_1)), (x_2=1, y_2=y(x_2), z_2=\varphi(y_2)) \right\}$\\
- \While{ $\Delta_t \geq \epsilon$ and  $k \leq K_{max}$}{
-  Compute the estimate of the H\"older constant $\mu$ according to (\ref{mu})\\
-	\For{$i=1$ \KwTo $k$}{
-	Compute the characteristic $R(i)$ according to (\ref{R})\\
-	}
-	Determine the number $t$, for which $R(t) = \max \left\{ R(i), 1 \leq i \leq k \right\}$
+The theory of algorithm convergence as well as some of its modifications are presented in \citet{Strongin2000}.
 
-	Compute the new trial point $x^{k+1}$ according to (\ref{xk1})
-
-	Perform a new trial at the point $x^{k+1}$
-
-	Add $(x^{k+1}, y^{k+1}, z^{k+1})$ to the set $\Omega_k$
-
-  $k = k + 1$\\
- }
- $l = \arg \min \left\{ z_i, 1 < i \leq k \right\}$\\
- $y_{min} = y_l$, \\ 
- \KwRet{$y_{min}$}
- \caption{Global search algorithm}\label{alg_GSA}
-\end{algorithm}
-
-The theory of algorithm convergence as well as some of its modifications are presented in \cite{Strongin2000}.
-
-\subsection{GSA for the problems with categorical parameters}
-\label{sec_mGSA}
+\subsection{GSA for the problems with categorical parameters}\label{sec_mGSA}
 
@@
 
-1.
-At the initialization stage, the values $(x^i = i, \nu^i=0, z^i)$, $ i, 0\leq i \leq S$, i.e., the boundary points of the intervals $(x_{i-1},x_i), 1<i\leq S$, are entered into the set $\Omega_k$ containing the search  information.
+\begin{enumerate}[1.]
+\item At the initialization stage, the values $(x^i = i, \nu^i=0, z^i)$, $ i, 0\leq i \leq S$, i.e., the boundary points of the intervals $(x_{i-1},x_i), 1<i\leq S$, are entered into the set $\Omega_k$ containing the search  information.
 After that, trials are carried out at arbitrary internal points (e.g., midpoints) of the intervals $(x_{i-1},x_i), 1<i\leq S$; the trial results are also added to the search information.
@@
 
-2.
-The computation of the estimate of the H\"older constant $\mu$ according to formula (\ref{mu}) must be performed only for intervals for which $\nu(x_i) = \nu(x_{i-1}) = 1$, i.e., when the values $z_i$ and $z_{i-1}$ are determined.
+\item The computation of the estimate of the H\"older constant $\mu$ according to formula (\ref{mu}) must be performed only for intervals for which $\nu(x_i) = \nu(x_{i-1}) = 1$, i.e., when the values $z_i$ and $z_{i-1}$ are determined.
 
-3. The computation of the characteristic of the search interval $(x_{i-1},x_i), 1<i\leq k$, is performed according to the following rules:
-\begin{gather}\label{R_int}
-R(i) = \Delta_i + \frac{(z_i-z_{i-1})^2}{(r\mu)^2\Delta_i}-2\frac{z_{i-1}+z_i}{r\mu}, \; \mathrm{if} \; \nu(x_i) = \nu(x_{i-1}) = 1, \nonumber \\ 
-R(i) = 2\Delta_i-4\frac{z_i}{r\mu}, \; \mathrm{if} \;  \nu(x_i) = 1, \nu(x_{i-1}) = 0, \nonumber \\ 
-R(i) = 2\Delta_i-4\frac{z_{i-1}}{r\mu}, \; \mathrm{if} \;  \nu(x_{i-1}) = 1, \nu(x_{i}) = 0. \nonumber
-\end{gather}
-
+\item The computation of the characteristic of the search interval $(x_{i-1},x_i), 1<i\leq k$, is performed according to the following rules:
+\begin{equation}\label{R_int}
+\begin{aligned}
+&R(i) = \Delta_i + \frac{(z_i-z_{i-1})^2}{(r\mu)^2\Delta_i}-2\frac{z_{i-1}+z_i}{r\mu},\\
+&\mathrm{if} \; \nu(x_i) = \nu(x_{i-1}) = 1,  \\ 
+&R(i) = 2\Delta_i-4\frac{z_i}{r\mu},\\
+&\mathrm{if} \; \nu(x_i) = 1, \nu(x_{i-1}) = 0,  \\ 
+&R(i) = 2\Delta_i-4\frac{z_{i-1}}{r\mu},\\
+&\mathrm{if} \; \nu(x_{i-1}) = 1, \nu(x_{i}) = 0. \nonumber
+\end{aligned}
+\end{equation}
 Note that the algorithm will not produce intervals whose both boundary points have undefined function values, i.e., $\nu(x_i) = \nu(x_{i-1}) = 0$.
 
-4. The new trial point $x^{k+1}$ will be calculated as follows:
-\begin{gather}\label{xk1_int}
-x^{k+1} = \frac{x_t+x_{t-1}}{2}- \frac{\mathrm{sign}(z_t-z_{t-1})}{2r} \left(\frac{\left|z_t-z_{t-1}\right|}{\mu}\right)^N, \; \mathrm{if} \; \nu(x_i) = \nu(x_{i-1}) = 1, \nonumber \\    
-x^{k+1} = \frac{x_t+x_{t-1}}{2} , \; \mathrm{if} \; \nu(x_i) \neq \nu(x_{i-1}). \nonumber
-\end{gather}
-
+\item The new trial point $x^{k+1}$ will be calculated as follows:
+\begin{equation}\label{xk1_int}
+\begin{aligned}
+&x^{k+1}=\frac{x_t+x_{t-1}}{2}-\\
+&-\frac{\mathrm{sign}(z_t-z_{t-1})}{2r}\left(\frac{\left|z_t-z_{t-1}\right|}{\mu}\right)^N,\\
+&\mathrm{if} \; \nu(x_i) = \nu(x_{i-1}) = 1,\\
+&x^{k+1} = \frac{x_t+x_{t-1}}{2},\\
+&\mathrm{if} \; \nu(x_i) \neq \nu(x_{i-1}). \nonumber
+\end{aligned}
+\end{equation}
 Here, $t$ is the number of the search interval that has the maximum characteristic.
 We note again that the characteristic $R(i)$ quantifies the possibility of finding a point $x^*$, which is the inverse image of the global optimizer $y^* = y(x^*)$, within the considered interval $(x_{i-1},x_i)$.
-
-%Новый текст на русском
-\textcolor[rgb]{1,0,0}{Условия сходимости Global Search Algorithm для задач с Categorical Parameters (GSA-CP) могут быть сформулированы в виде следующего утверждения.}
-
-\textbf{Утверждение} (сonvergence conditions)
-
-Пусть функции $ \varphi(y,u_s), y \in D, s = 1,...,S,$ form (\ref{problem_is}) удовлетворяют условию Липшица с соответствующими константами $L_s, s = 1,...,S$, и $L = \max\{L_s\}$.
-И пусть точка $\overline{x}$ be a limit point of the sequence $\left\{x^k\right\} $, полученной при минимизации функции $f(x)$ из (\ref{f}) с помощью алгоритма GSA-CP. Then:
-\begin{enumerate}
-	\item 
-	если одноверменно с $\overline{x}$ существует другая предельная точка $x'$ последовательности $\left\{x^k\right\}$, тогда $f(\overline{x}) = f(x')$.
-	\item
-	$z^k=f(x^k) \geq f(\overline{x})$ при любом $k \geq 0$.
-	\item
-	Если на некотором шаге поиска величина $\mu$ from (\ref{mu}) satisfies the condition
-\begin{equation}\label{217}
-  r \mu > 2^{3-1/N} L \sqrt{N+3},
-\end{equation}
-then $\overline{x}$ is a global minimizer of the function $f(x), x\in[0,S]$ and any global minimizer $x^\ast$ is also a limit point of the sequence $\left\{x^k\right\}$.
 \end{enumerate}
 
-\textbf{Proof}.
-Это утверждение доказывается аналогично теореме о сходимости index method of global optimization from \cite{Strongin2000}, принимая во внимание взаимосвясь (\ref{f}) функций  $f(x)$ и $\varphi(y,u_s)$, а также правила (\ref{index}), задающие индексы у точек испытаний.
-
-\section{Numerical experiments}
-\label{sec_exp}
-
-In this section, we present experimental results that demonstrate the effectiveness of the developed framework for a number of HPO problems.
-First, we describe in detail the setup of the computational experiment.
-Then, the results of comparing the iOpt framework with several well-known frameworks widely used for hyperparameter tuning in machine learning methods are presented.
-The section closes with conclusions and a discussion of the results.
-
-\subsection{Experimental setup}
-
-\textbf{Datasets.} 
-We used for our experiments fifteen datasets from the UCI repository \cite{UCI}, which covers a large number of areas.
-Datasets from this repository are quite often used to test the correct operation of hyperparameter tuning algorithms \cite{Joy2020,Wang2021,Wu2023}.
-We selected the datasets to ensure that the data contained in them were described by both numerical and categorical attributes.
-Note that most of the datasets were unbalanced and the data presented in the datasets were adapted for the classification task.
-The characteristics of the datasets used are presented in Table \ref{tab:1}.
-
-\begin{table}[ht]
+\begin{table*}
+\centering
 \caption{The datasets used in the experiments}
 \label{tab:1}
-\resizebox{\textwidth}{!}{%
+\resizebox{0.8\textwidth}{!}{%
 \begin{tabular}{ccccccc}
@@
 }
-\end{table}
+\end{table*}
 
-\textbf{Classification algorithms.} We performed hyperparameter tuning for three types of classification algorithms on all datasets.
+%Новый текст на русском
+\textcolor[rgb]{1,0,0}{
+Условия сходимости Global Search Algorithm для задач с Categorical Parameters (GSA-CP) могут быть сформулированы в виде следующего утверждения.
+\begin{proposition}
+Пусть функции $ \varphi(y,u_s), y \in D, s = 1,...,S,$ form (\ref{problem_is}) удовлетворяют условию Липшица с соответствующими константами $L_s, s = 1,...,S$, и $L = \max\{L_s\}$.
+И пусть точка $\overline{x}$ be a limit point of the sequence $\left\{x^k\right\} $, полученной при минимизации функции $f(x)$ из (\ref{f}) с помощью алгоритма GSA-CP. Then:
 \begin{enumerate}
-	\item XGBoost is one of the tree ensemble construction methods, which is an efficient implementation of gradient boosting over decision trees \cite{XGBoost2016}.
-The main advantages of the algorithm include a smart penalty mechanism, proportional reduction of leaf nodes, and automatic feature selection.
-It is known that the performance of an algorithm and its predictive ability depend significantly on the correct tuning of hyperparameters.
-The implementation of the method is based on the widely used XGBoost library.
- \item Multi-Layer Perceptron (MLP), a fully connected artificial neural network, is a supervised learning algorithm that forms a function $f : R^m \to R^o$, where $m$ is the number of input attributes, and $o$ is the output dimensionality of the network \cite{GARDNER19982627}.
-Using this method, it is possible to form the function f as a universal nonlinear approximator for the classification and regression problem.
-In the framework of this study, we will consider the multilayer perceptron for the task of attributing an object to a particular class.
- \item Support Vector Classification (SVC) is a support vector method for the classification problem \cite{SVC}.
-The main purpose of this algorithm is to find the equation of a separating hyperplane $\omega_0 + \sum_{i=1}^n \omega_i x_i=0$ in the space $R^n$ that separates classes in some optimal way.
-To improve the predictive ability of the algorithm, a method for constructing a nonlinear classifier was proposed by switching from scalar products to arbitrary kernels, which allows the construction of nonlinear separators.
-An implementation of the SVC method from the scikit-learn library was used to investigate the capabilities of HPO algorithms.
+	\item 
+	если одноверменно с $\overline{x}$ существует другая предельная точка $x'$ последовательности $\left\{x^k\right\}$, тогда $f(\overline{x}) = f(x')$.
+	\item
+	$z^k=f(x^k) \geq f(\overline{x})$ при любом $k \geq 0$.
+	\item
+	Если на некотором шаге поиска величина $\mu$ from (\ref{mu}) satisfies the condition
+\begin{equation}\label{217}
+  r \mu > 2^{3-1/N} L \sqrt{N+3},
+\end{equation}
+then $\overline{x}$ is a global minimizer of the function $f(x), x\in[0,S]$ and any global minimizer $x^\ast$ is also a limit point of the sequence $\left\{x^k\right\}$.
+\end{enumerate}
+\end{proposition}
+\begin{proof}
+Это утверждение доказывается аналогично теореме о сходимости index method of global optimization from \citet{Strongin2000}, принимая во внимание взаимосвясь (\ref{f}) функций  $f(x)$ и $\varphi(y,u_s)$, а также правила (\ref{index}), задающие индексы у точек испытаний.
+\end{proof}
+}
 
-\end{enumerate}
+\section{Numerical experiments}\label{sec_exp}
+In this section, we present experimental results that demonstrate the effectiveness of the developed framework for a number of HPO problems.
+First, we describe in detail the setup of the computational experiment.
+Then, the results of comparing the iOpt framework with several well-known frameworks widely used for hyperparameter tuning in machine learning methods are presented.
+The section closes with conclusions and a discussion of the results.
 
-The tunable hyperparameters of the above algorithms and their search region are listed in Table \ref{tab:2}.
+\subsection{Experimental setup}
 
-\begin{table}[ht]
+\textbf{Datasets.} We used for our experiments fifteen datasets from the UCI repository \citep{UCI}, which covers a large number of areas.
+Datasets from this repository are quite often used to test the correct operation of hyperparameter tuning algorithms \citep{Joy2020,Wang2021,Wu2023}.
+We selected the datasets to ensure that the data contained in them were described by both numerical and categorical attributes.
+Note that most of the datasets were unbalanced and the data presented in the datasets were adapted for the classification task.
+The characteristics of the datasets used are presented in Table \ref{tab:1}.
+
+\begin{table*}
+\centering
 \caption{Tunable hyperparameters of machine learning algorithms.}
 \label{tab:2}
+\resizebox{0.9\textwidth}{!}{%
 \begin{tabular}{cccc}
@@
                          & learning\_rate       & Float         & {[}0.001, 0.1{]}                                                       \\ \hline
-\multirow{3}{*}{\begin{tabular}[c]{@{}l@{}}Support \\ Vector   \\ Classification \\ (SVC)\end{tabular}} &
+\multirow{3}{*}{\begin{tabular}[c]{@{}l@{}}Support Vector \\ Classification (SVC)\end{tabular}} &
   Gamma &
@@
                          & C                    & Int           & {[}$10^1$,$10^{10}${]}                     \\
-                         & Kernel               & Categorial    & \begin{tabular}[c]{@{}l@{}}\{‘poly’, ‘rbf’,\\ ‘sigmoid’\}\end{tabular} \\ \hline
+                         & Kernel               & Categorial    & \begin{tabular}[c]{@{}l@{}}\{'poly', 'rbf','sigmoid'\}\end{tabular} \\ \hline
 \multirow{4}{*}{\begin{tabular}[c]{@{}l@{}}Multi-layer   \\ Perceptron \\ (MLP)\end{tabular}} &
@@
   Categorial &
-  \begin{tabular}[c]{@{}l@{}}\{'identity', \\ 'logistic', \\ 'tanh', 'relu'\}\end{tabular} \\
+  \begin{tabular}[c]{@{}l@{}}\{'identity', 'logistic', 'tanh', 'relu'\}\end{tabular} \\
                          & hidden\_layer\_sizes & Int           & {[}2, 30{]}                                                            \\
                          & max\_iter            & Int           & {[}100, 700{]}                                                         \\
-                         & Solver               & Categorial    & \begin{tabular}[c]{@{}l@{}}\{'lbfgs', 'sgd', \\ 'adam'\}\end{tabular}  \\ \hline
+                         & Solver               & Categorial    & \begin{tabular}[c]{@{}l@{}}\{'lbfgs', 'sgd', 'adam'\}\end{tabular}  \\ \hline
 \end{tabular}
-\end{table}
+}
+\end{table*}
+
+\textbf{Classification algorithms.} We performed hyperparameter tuning for three types of classification algorithms on all datasets.
+\begin{enumerate}
+\item XGBoost is one of the tree ensemble construction methods, which is an efficient implementation of gradient boosting over decision trees \citep{XGBoost2016}.
+The main advantages of the algorithm include a smart penalty mechanism, proportional reduction of leaf nodes, and automatic feature selection.
+It is known that the performance of an algorithm and its predictive ability depend significantly on the correct tuning of hyperparameters.
+The implementation of the method is based on the widely used XGBoost library.
+\item Multi-Layer Perceptron (MLP), a fully connected artificial neural network, is a supervised learning algorithm that forms a function $f : R^m \to R^o$, where $m$ is the number of input attributes, and $o$ is the output dimensionality of the network \citep{GARDNER19982627}.
+Using this method, it is possible to form the function f as a universal nonlinear approximator for the classification and regression problem.
+In the framework of this study, we will consider the multilayer perceptron for the task of attributing an object to a particular class.
+\item Support Vector Classification (SVC) is a support vector method for the classification problem \citep{SVC}.
+The main purpose of this algorithm is to find the equation of a separating hyperplane $\omega_0 + \sum_{i=1}^n \omega_i x_i=0$ in the space $R^n$ that separates classes in some optimal way.
+To improve the predictive ability of the algorithm, a method for constructing a nonlinear classifier was proposed by switching from scalar products to arbitrary kernels, which allows the construction of nonlinear separators.
+An implementation of the SVC method from the scikit-learn library was used to investigate the capabilities of HPO algorithms.
+\end{enumerate}
+
+The tunable hyperparameters of the above algorithms and their search region are listed in Table \ref{tab:2}.
 
@@
 \item Accuracy is a metric describing the overall prediction accuracy of the method across all classes.
-It is most visible in the case when each class is equally important and the dataset shows a balance of classes \cite{Metric1}.
+It is most visible in the case when each class is equally important and the dataset shows a balance of classes \citep{Metric1}.
 This metric has been applied when working with the datasets Optical Recognition of Handwritten Digits, Bank Marketing, CNAE9, Statlog Segmentation, Semeion Handwritten Digit;
- \item Binary F1 -- F1 score is adapted for binary classification tasks, with the predominance of one class over another in the dataset \cite{Chicco2020}.
+\item Binary F1 -- F1 score is adapted for binary classification tasks, with the predominance of one class over another in the dataset \citep{Chicco2020}.
 This metric is suitable for analyzing Breast Cancer Wisconsin, Credit Approval, Parkinsons, Banknote authentication, Wilt datasets;
- \item Macro F1 -- F1 score is suitable for tasks with a large number of unbalanced classes \cite{Metric2}.
+\item Macro F1 -- F1 score is suitable for tasks with a large number of unbalanced classes \citep{Metric2}.
 The datasets Protein Localization Sites (Ecoli), Balance Scale Weight \& Distance, Zoo, Car Evaluation, Turkiye Student Evaluation have this particular feature.
-
 \end{itemize}
 
-\textbf{HPO algorithms.} The hyperparameters of machine learning methods were tuned using the iOpt framework developed by us, and also using the well-known frameworks Optuna \cite{optuna} and HyperOpt \cite{hyperopt}.
+\textbf{HPO algorithms.} The hyperparameters of machine learning methods were tuned using the iOpt framework developed by us, and also using the well-known frameworks Optuna \citep{optuna} and HyperOpt \citep{hyperopt}.
 
@@
 In the study, we used the following hyperparameter search options:
+
 \begin{enumerate}
 	\item Optuna(random) will perform a random search. For each hyperparameter, a distribution is specified, from which its value is selected.
-	\item Optuna(tpe) is a sampler based on the Tree-structured Parzen Estimator algorithm \cite{NIPS2011_86e8f7ab}.
+	\item Optuna(tpe) is a sampler based on the Tree-structured Parzen Estimator algorithm \citep{NIPS2011}.
 On each trial, for each parameter, TPE fits one Gaussian Mixture Model (GMM) $l(x)$ to the set of parameter values associated with the best objective values, and another GMM $g(x)$ to the remaining parameter values.
 It chooses the parameter value x that maximizes the ratio $l(x)/g(x)$.
-	\item Optuna(cmaes) is an algorithm based on CMA-ES (Lightweight Covariance Matrix Adaptation Evolution Strategy) \cite{cmaes2015}.
+	\item Optuna(cmaes) is an algorithm based on CMA-ES (Lightweight Covariance Matrix Adaptation Evolution Strategy) \citep{cmaes2015}.
 The CMA-ES is a stochastic method for real-parameter optimization of non-linear, non-convex functions.
 The algorithm derived from intuitive concepts and from requirements of non-linear, non-convex search in continuous domain.
-	\item Optuna (nsgaii) is a sampler based on the Nondominated Sorting Genetic Algorithm II,  which is a well known, fast and elitist multi-objective genetic algorithm \cite{nsgaii2013}.
-
+	\item Optuna (nsgaii) is a sampler based on the Nondominated Sorting Genetic Algorithm II,  which is a well known, fast and elitist multi-objective genetic algorithm \citep{nsgaii2013}.
 \end{enumerate}
@@
 There are currently three algorithms implemented in Hyperopt: Random Search, Tree of Parzen Estimators (TPE) and Adaptive TPE.
-We used the computation variant with the TPE algorithm implementation -- Hyperopt(tpe) \cite{NIPS2011_86e8f7ab}.
+We used the computation variant with the TPE algorithm implementation -- Hyperopt(tpe) \citep{NIPS2011}.
 
 When using the iOpt framework, we applied a variant of computation in which the available resource was divided into portions by the number of trials.
-Most of the trials (85\%) were performed using the global search algorithm (see Section \ref{sec_iOpt}), while in the remaining trials we used the local method to refine the solution found \cite{Kelley}.
+Most of the trials (85\%) were performed using the global search algorithm (see Section \ref{sec_iOpt}), while in the remaining trials we used the local method to refine the solution found \citep{Kelley}.
 
@@
 
-In order to check the statistical significance of the experimental results, the Friedman test was applied.
-An implementation of the corresponding test from the SciPy library was used.
-The target metric values obtained by Optuna, Hyperopt and iOpt framework runs were used as input data.
-The test was carried out separately for each of the algorithms for which hyperparameters were tuned: SVC, XGBoost, MLP.
-The null hypothesis about the consistent work of the compared methods is rejected at significance level $p=0.01$.
-Note that in some cases the $p$-value was undefined (for example, when analyzing the ``Banknote'' data set using the MLP method and ``Zoo'' using the SVC method), which is due to the equality of the target metric for all optimization runs.
-
-The values of the classification metrics and their standard deviation for the SVC, XGBoost and MLP algorithms are presented in Table \ref{tab:res}. 
-The values corresponding to the best result on average are highlighted in bold. The values corresponding to the guaranteed best result, i.e. the best result for the worst case, are highlighted in red.
-
-It is obvious from Table \ref{tab:res} that the iOpt framework is comparable in terms of quality to Hyperopt and Optuna.
-When using various methods for constructing classification models and different datasets based on the average metric values, iOpt allows us to obtain the best values in 15 models out of 42.
-It is worth noting that algorithms which involve random variables in one way or another do not guarantee the best values of the objective function for one run.
-To obtain good results, a stochastic algorithm must be run several times.
-If one focuses on a guaranteed result, iOpt works better in more than half of the cases (23 out of 42 models).
-iOpt shows the best results when tuning hyperparameters for the SVC model.
-For most of the remaining models, the value of quality metrics differs only slightly.
-
-%Figures \ref{fig_res_1} and \ref{fig_res_2} show the values of the classification metrics of the SVC algorithm in the worst and average case, respectively.
-As can be seen, when optimizing the hyperparameters of the SVC algorithm on the Banknote, CNAE-9, Statlog Segmentation, and Wilt datasets, the iOpt framework has demonstrated the best results.
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.85\textwidth]{SVC_1.pdf}
-%\caption{Values of the optimized metrics for the SVC algorithm, obtained in the worst case} \label{fig_res_1}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.85\textwidth]{SVC_2_avg.pdf}
-%\caption{Values of optimized metrics for the SVC algorithm averaged over 10 runs} \label{fig_res_2}
-%\end{figure}
-%
-%
-%On the CarEvaluation, Credit Approval, Ecoli, Parkinsons, and Semeion datasets, the best values of the metrics on average were obtained with the Optuna (tpe).
-On the datasets Balance, Breast Cancer, Digits, on average the Optuna(cmaes) performed well.
-The Hyperopt(tpe) performed better than the others on the Bank-marketing dataset.
-In the case of the Zoo dataset, all optimization algorithms showed the same results.
-%
-%It is especially worth noting that the algorithms implemented in both Optuna and Hyperopt contain random parameters; as a consequence, the best values of the objective metric are not guaranteed for a single run.
-%
-%This fact is illustrated by Fig.
-\ref{fig_res_1}.
-The iOpt framework additionally performed better on the Bank-marketing, Banknote, CNAE-9, Statlog Segmentation, and Wilt datasets.
-Thus on 9 out of 14 datasets the iOpt framework showed better values of the objective metric compared to the Optuna and Hyperopt frameworks.
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.85\textwidth]{XGBoost_1.pdf}
-%\caption{Values of the optimized metrics for the XGBoost algorithm, obtained in the worst case} \label{fig_res_3}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.85\textwidth]{XGBoost_2_avg.pdf}
-%\caption{Values of optimized metrics for the XGBoost algorithm averaged over 10 runs} \label{fig_res_4}
-%\end{figure}
-%
-%Figures \ref{fig_res_3} and \ref{fig_res_4} show the value of classification metrics of the XGBoost algorithm in the worst and average case, respectively.
-In the case of optimizing the hyperparameters  of the XGBoost algorithm, the iOpt framework showed better results on the CarEvaluation, CNAE-9, Credit Approval, Ecoli, Parkinsons, Wilt, Zoo datasets.
-%
-%On average, when using 10 runs, the best performance of the Optuna(tpe) framework was observed on the Balance, Bank-marketing, Banknote, Breast Cancer, CarEvaluation, Credit Approval, Digits, Ecoli, Semeion, Statlog Segmentation, Wilt datasets.
-The Hyperopt(tpe) showed good results on the CNAE-9 and Parkinsons datasets on average for 10 runs.
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.85\textwidth]{MLP_1.pdf}
-%\caption{Values of the optimized metrics for the XGBoost algorithm, obtained in the worst case} \label{fig_res_5}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.85\textwidth]{MLP_2_avg.pdf}
-%\caption{Values of optimized metrics for the XGBoost algorithm averaged over 10 runs} \label{fig_res_6}
-%\end{figure}
-%
-%Figures \ref{fig_res_5} and \ref{fig_res_6} show the values of classification metrics of the MLP algorithm.
-The iOpt framework performed better on the Balance, Bank-marketing, Banknote, Breast Cancer, Credit Approval, Wilt, Zoo datasets.
-Considering the average values of metrics on 10 runs of hyperparameter optimization, the Optuna framework performed better on the Breast Cancer, CNAE-9, Credit Approval, Digits, Ecoli, Parkinsons, Semeion datasets, and the Hyperopt on the Statlog Segmentation, Wilt, CarEvaluation datasets.
-%
-
-%Thus, we can conclude that the iOpt framework shows comparable results with the Optuna and Hyperopt frameworks in terms of performance quality.
-It should be reminded that the algorithms implemented within the iOpt framework are deterministic.
-As a consequence, iOpt, unlike Optuna and Hyperopt, only needs to be run once to achieve the best values of the quality metric.
-
 % Please add the following required packages to your document preamble:
@@
 % If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
-\begin{table}
+\begin{table*}
 \centering
@@
 }
-\end{table}
+\end{table*}
 
-Next, we consider the performance of the iOpt, Optuna, and Hyperopt frameworks. Table \ref{tab:time} shows the running time of the frameworks for tuning the hyperparameters.
-Since the random search did not show good results in terms of the evaluation metrics, we did not include the corresponding column in the table. 
-A single optimization run was performed with a maximum of 200 search trials.
-The values corresponding to the best result on average are highlighted in bold.
-It can be clearly seen that even in the case of a single run iOpt shows better performance on many datasets compared to Optuna and Hyperopt (29 out of 42 models).
+In order to check the statistical significance of the experimental results, the Friedman test was applied.
+An implementation of the corresponding test from the SciPy library was used.
+The target metric values obtained by Optuna, Hyperopt and iOpt framework runs were used as input data.
+The test was carried out separately for each of the algorithms for which hyperparameters were tuned: SVC, XGBoost, MLP.
+The null hypothesis about the consistent work of the compared methods is rejected at significance level $p=0.01$.
+Note that in some cases the $p$-value was undefined (for example, when analyzing the ``Banknote'' data set using the MLP method and ``Zoo'' using the SVC method), which is due to the equality of the target metric for all optimization runs.
 
-To obtain a good solution, the methods from Optuna and Hyperopt should be restarted several times.
-If one focuses on the total operating time, the performance gain of iOpt is even greater.
-As an example, Figure \ref{fig8} shows a comparison of the hyperparameter tuning times for iOpt, Optuna and Hyperopt on several datasets and different classification models.
-For iOpt, the time is shown as a blue dot.
-For Optuna and Hyperopt, the box plots are shown.
+The values of the classification metrics and their standard deviation for the SVC, XGBoost and MLP algorithms are presented in Table \ref{tab:res}. 
+The values corresponding to the best result on average are highlighted in bold. The values corresponding to the guaranteed best result, i.e. the best result for the worst case, are highlighted in red.
+
+It is obvious from Table \ref{tab:res} that the iOpt framework is comparable in terms of quality to Hyperopt and Optuna.
+When using various methods for constructing classification models and different datasets based on the average metric values, iOpt allows us to obtain the best values in 15 models out of 42.
+It is worth noting that algorithms which involve random variables in one way or another do not guarantee the best values of the objective function for one run.
+To obtain good results, a stochastic algorithm must be run several times.
+If one focuses on a guaranteed result, iOpt works better in more than half of the cases (23 out of 42 models).
+iOpt shows the best results when tuning hyperparameters for the SVC model.
+For most of the remaining models, the value of quality metrics differs only slightly.
 
@@
 % \usepackage{graphicx}
-\begin{table}
+\begin{table*}
 \centering
@@
 }
-\end{table}
+\end{table*}
 
-%\begin{table}[ht]
-%\caption{Time of hyperparameter tuning for a single run of algorithms.}
-%\label{tab:3}
-%\begin{tabular}{lllllll}
-%\hline
-%Dataset &
-%  \begin{tabular}[c]{@{}l@{}}Optuna \\ random\end{tabular} &
-%  \begin{tabular}[c]{@{}l@{}}Optuna \\ tpe\end{tabular} &
-%  \begin{tabular}[c]{@{}l@{}}Optuna \\ cmaes\end{tabular} &
-%  \begin{tabular}[c]{@{}l@{}}Optuna \\ nsgaii\end{tabular} &
-%  \begin{tabular}[c]{@{}l@{}}Hyperopt\\ tpe\end{tabular} &
-%  iOpt \\ \hline
-%Balance                                                         & 10.8   & 9.2    & 9.0    & 10.6   & 10.3   & \textbf{7.7}    \\
-%\begin{tabular}[c]{@{}l@{}}Bank \\ Marketing\end{tabular}       & 2888.5 & 3406.5 & \textbf{2704.7} & 3078.6 & 3190.3 & 2963.3 \\
-%Banknote                                                        & 31.9   & \textbf{12.0}   & 17.7   & 22.1   & 21.7   & 18.4   \\
-%\begin{tabular}[c]{@{}l@{}}Breast \\ Cancer\end{tabular}        & 9.3    & 7.6    & 6.3    & 8.1    & 8.5    & \textbf{6.0}    \\
-%CarEvaluation                                                   & 50.9   & 63.3   & 51.0   & 53.1   & 69.5   & \textbf{45.2}   \\
-%CNAE9                                                           & 295.6  & 207.6  & 318.4  & 262.2  & 253.5  & \textbf{144.9}  \\
-%\begin{tabular}[c]{@{}l@{}}Credit \\ Approval\end{tabular}      & 14.2   & 12.6   & 13.4   & 13.4   & 15.8   & \textbf{10.8}   \\
-%Digits                                                          & 132.8  & \textbf{57.5}   & 82.1   & 111.6  & 92.4   & 62.5   \\
-%Ecoli                                                           & 5.5    & 6.3    & 4.0    & 5.4    & 6.4    & \textbf{3.6}    \\
-%Parkinsons                                                      & 3.5    & 5.0    & 3.6    & 3.8    & 4.4    & \textbf{2.9}    \\
-%Semeion                                                         & 218.4  & 119.5  & 157.9  & 135.7  & 193.6  & \textbf{114.2}  \\
-%\begin{tabular}[c]{@{}l@{}}Statlog \\ Segmentation\end{tabular} & 156.1  & \textbf{55.4}   & 87.1   & 159.2  & 91.6   & 66.5   \\
-%Wilt                                                            & 70.0   & 75.8   & 68.5   & 67.4   & 78.2   & \textbf{60.9}   \\
-%Zoo                                                             & 3.0    & 3.9    & 3.0    & 3.1    & 4.2    & \textbf{1.8}    \\ \hline
-%\end{tabular}
-%\end{table}
+Next, we consider the performance of the iOpt, Optuna, and Hyperopt frameworks. Table \ref{tab:time} shows the running time of the frameworks for tuning the hyperparameters.
+Since the random search did not show good results in terms of the evaluation metrics, we did not include the corresponding column in the table. 
+A single optimization run was performed with a maximum of 200 search trials.
+The values corresponding to the best result on average are highlighted in bold.
+It can be clearly seen that even in the case of a single run iOpt shows better performance on many datasets compared to Optuna and Hyperopt (29 out of 42 models).
 
-\begin{figure}
+To obtain a good solution, the methods from Optuna and Hyperopt should be restarted several times.
+If one focuses on the total operating time, the performance gain of iOpt is even greater.
+As an example, Figure \ref{fig8} shows a comparison of the hyperparameter tuning times for iOpt, Optuna and Hyperopt on several datasets and different classification models.
+For iOpt, the time is shown as a blue dot.
+For Optuna and Hyperopt, the box plots are shown.
+
+\begin{figure*}
 \centering
@@
 \caption{Time of hyperparameters tuning.} \label{fig8}
-\end{figure}
+\end{figure*}
 
-\section{Conclusion}
+\subsection{Results on surrogate models}
+
+Дополнительная серия экспериментов по сравнению iOpt, Optuna и HyperOpt была проведена с использованием фреймворка JAHS-Bench-201 [ссылка].
+JAHS-Bench предоставляет удобный доступ к набору задач для определения оптимального сочетания гиперпараметров и архитектуры neural network.
+Данный инструмент для сравнения HPO-алгоритмов содержит обширный набор данных о качестве и производительности нейронных сетей, состоящий примерно из 161 million data points, каждая из которых состоит из 20 показателей.
+
+\begin{table*}
+\centering
+\caption{Values of optimized metrics for the algorithms averaged over 10 runs (surrogate models)}
+\label{tab:met-surr}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{ccccccc}
+\hline
+\textbf{}                      & \textbf{iOpt}   & \textbf{Optuna(random)} & \textbf{Optuna(tpe)}     & \textbf{Optuna(cmaes)}   & \textbf{Optuna(nsgaii)} & \textbf{Hyperopt(tpe)} \\ \hline
+\textbf{cifar10}               & {\color[HTML]{FE0000} \textbf{88.394} } & 88.144 $\pm$ 0.21\%         & 88.287 $\pm$ 0.21\%          & 88.218 $\pm$ 0.15\%          & 88.063 $\pm$ 0.27\%         & 88.316 $\pm$ 0.11\%        \\
+\textbf{fashion\_mnist}        & 94.797          & 94.677 $\pm$ 0.18\%         & 94.837 $\pm$ 0.13\%          & {\color[HTML]{FE0000} \textbf{94.952 $\pm$ 0.08\%} } & 94.631 $\pm$ 0.35\%         & 94.759 $\pm$ 0.19\%        \\
+\textbf{colorectal\_histology} & {\color[HTML]{FE0000} \textbf{94.447} } & 94.013 $\pm$ 0.33\%         & \textbf{94.447 $\pm$ 0.17\%} & 94.194 $\pm$ 1.18\%          & 94.150 $\pm$ 0.36\%         & 94.233 $\pm$ 0.17\%        \\ \hline
+\end{tabular}%
+}
+\end{table*}
+
+Особенностью JAHS-Bench является использование surrogate models fit on performance data, covering different optimization problems (single/multiple objectives), что предоставляет возможность варьировать архитектуру нейросети (выбирая один из имеющихся вариантов), а также ее гиперпараметры.
+При этом использование surrogate models позволяет оценить качество работы обученного классификатора при конкретном наборе гиперпараметров за доли секунды вместо нескольких часов, которые были бы затрачены на обучение реальной нейросети.
+В качестве основы для суррогатных моделей в JAHS-Bench используются три популярных датасета -- CIFAR-10, Colorectal-Histology, Fashion MNIST -- применяемых для задач классификации.
+
+В проведенных экспериментах настройка гиперпараметров с помощью фреймворков iOpt, Optuna и HyperOpt выполнялась на фиксированной архитектуре нейросети, предложенной в JAHS-Bench качестве архитектуры по умолчанию.
+Для настройки были выбраны традиционно наиболее значимые (с точки зрения исследователя) гиперпараметры: категориальный параметр выбора функции активации (принимающий три возможных значения ReLU, Mish, Hardswish), а также непрерывные параметры Learning Rate and Weight Decay в диапазонах $[10^{-3},1]$ and $[10^{-5},10^{-2}]$ соответственно.
+
+В качестве целевой функции использовалась метрика validation accuracy, вычисленная после 200 эпох обучения.
+Точность поиска в критерии остановки всех методов была установлена равной 0.01, а максимальное число испытаний – равным 200.
+Для алгоритмов из Optuna и HyperOpt достигнутые показатели усреднялись по 10 запускам.
+Результаты сравнения представлены в таблице \ref{tab:met-surr}.
+Как и ранее, the values corresponding to the best result on average are highlighted in bold and the values corresponding to the best result for the worst case are highlighted in red.
+
+Результаты показывают превосходство iOpt на датасетах CIFAR-10 и Colorectal-Histology.
+При использовании датасета Fashion MNIST фреймворк iOpt превосходит HyperOpt, но уступает Optuna.
+Отметим, что при использовании суррогатных моделей время вычисления целевой функции составляет доли секунды, поэтому результаты, связанные со сравнением времени работы алгоритмов, для этого эксперимента не приводятся.
+
+\section{Conclusion}\label{sec13}
+
 In this paper, we propose a hyperparameter tuning method that is based on the ideas of Lipschitz global optimization.
@@
 
-\section*{CRediT authorship contribution statement}
+%\backmatter
+\bmhead{Supplementary information}
 
-\textbf{Konstantin Barkalov:} Supervision, Conceptualization, Writing - original draft.
-\textbf{Denis Karchkov:} Software, Data curation.
-\textbf{Evgeny Kozinov:} Methodology, Validation, Writing - original draft.
-\textbf{Ilya Lebedev:} Software, Investigation, Writing - original draft.
-\textbf{Denis Rodionov:} Software, Investigation.
-\textbf{Marina Usova:} Visualization.
+The software that implements all of the described methods is available in the open repository \url{https://github.com/aimclub/iOpt/}.
+Python scripts that reproduce the results of the experiments described in this paper are available in the benchmarks folder.
 
-\section*{Declaration of competing interest}
+\bmhead{Acknowledgements}
+
+The research was carried out within the state assignment of Ministry of Science and Higher Education of the Russian Federation (project No FSER-2024-0004).
+
+%\section*{CRediT authorship contribution statement}
+
+%\textbf{Konstantin Barkalov:} Supervision, Conceptualization, Writing - original draft.
+%\textbf{Denis Karchkov:} Software, Data curation.
+%\textbf{Evgeny Kozinov:} Methodology, Validation, Writing - original draft.
+%\textbf{Ilya Lebedev:} Software, Investigation, Writing - original draft.
+%\textbf{Denis Rodionov:} Software, Investigation.
+%\textbf{Marina Usova:} Visualization.
+
+\section*{Declarations}
 
@@
 
-\section*{Code availability}
-The software that implements all of the described methods is available in the open repository \url{https://github.com/aimclub/iOpt/}.
-Python scripts that reproduce the results of the experiments described in this paper are available in the benchmarks folder.
+%%=============================================%%
+%% For submissions to Nature Portfolio Journals %%
+%% please use the heading ``Extended Data''.   %%
+%%=============================================%%
 
-\section*{Acknowledgements}
-This work was supported by the Analytical Center for the Government of the Russian Federation (IGK 000000D730321P5Q0002), agreement No. 70-2021-00141.
+%%=============================================================%%
+%% Sample for another appendix section			       %%
+%%=============================================================%%
 
-%% The Appendices part is started with the command \appendix;
-%% appendix sections are then done as normal sections
-%% \appendix
+%% \section{Example of another appendix section}\label{secA2}%
+%% Appendices may be used for helpful, supporting or essential material that would otherwise 
+%% clutter, break up or be distracting to the text. Appendices can consist of sections, figures, 
+%% tables and equations etc.
 
-%% \section{}
-%% \label{}
+%\end{appendices}
 
-%% If you have bibdatabase file and want bibtex to generate the
-%% bibitems, please use
-%%
-\bibliographystyle{elsarticle-num} 
-\bibliography{bibliography}
+%%===========================================================================================%%
+%% If you are submitting to one of the Nature Portfolio journals, using the eJP submission   %%
+%% system, please include the references within the manuscript file itself. You may do this  %%
+%% by copying the reference list from your .bbl file, paste it into the main manuscript .tex %%
+%% file, and delete the associated \verb+\bibliography+ commands.                            %%
+%%===========================================================================================%%
 
-%% else use the following coding to input the bibitems directly in the
-%% TeX file.
+%\bibliographystyle{abbrvnat}
+\bibliography{sn-bibliography}
+% common bib file
+%% if required, the content of .bbl file can be included here once bbl is generated
+%%\input sn-article.bbl
 
-%\begin{thebibliography}{00}
-
-%% \bibitem{label}
-%% Text of bibliographic item
-
-%\bibitem{}
-
-%\end{thebibliography}
 \end{document}
-\endinput
-%%
-%% End of file `elsarticle-template-num.tex'.

### 2024-03-18 2024/ITMO paper/sn-article.tex
Declarations ans Statements, References

@@
 %\documentclass[sn-mathphys-num]{sn-jnl}% Math and Physical Sciences Numbered Reference Style 
-\documentclass[iicol]{sn-jnl}
+\documentclass[lineno,iicol]{sn-jnl}
 %%\documentclass[sn-mathphys-ay]{sn-jnl}% Math and Physical Sciences Author Year Reference Style
@@
 
+%\usepackage[pagewise]{lineno}
+
 \usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
 %\setcitestyle{authoryear,open={(},close={)}} %Citation-related commands
@@
 %\backmatter
-\bmhead{Supplementary information}
+%\bmhead{Supplementary information}
 
-The software that implements all of the described methods is available in the open repository \url{https://github.com/aimclub/iOpt/}.
-Python scripts that reproduce the results of the experiments described in this paper are available in the benchmarks folder.
+%The software that implements all of the described methods is available in the open repository \url{https://github.com/aimclub/iOpt/}.
+%Python scripts that reproduce the results of the experiments described in this paper are available in the benchmarks folder.
 
-\bmhead{Acknowledgements}
+%\bmhead{Acknowledgements}
 
-The research was carried out within the state assignment of Ministry of Science and Higher Education of the Russian Federation (project No FSER-2024-0004).
+%The research was carried out within the state assignment of Ministry of Science and Higher Education of the Russian Federation (project No FSER-2024-0004).
 
@@
 
-\section*{Declarations}
+\section*{Statements and Declarations}
 
-The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
+\textbf{Funding} The research was carried out within the state assignment of Ministry of Science and Higher Education of the Russian Federation (project No FSER-2024-0004).
+
+\noindent
+\textbf{Competing Interests} The authors have no relevant financial or non-financial interests to disclose.
+
+\noindent
+\textbf{Data Availability} 
+The datasets analyzed during the current study are available in the UCI Machine Learning Repository, \url{https://archive.ics.uci.edu/}. 
+The software that implements all of the described methods is available in the open repository \url{https://github.com/aimclub/iOpt/}.
+Python scripts that reproduce the results of the experiments described in this paper are available in the benchmarks folder.
+
+%The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
+
+\section*{Compliance with Ethical Standards}
+
+\textbf{Conflict of interest} The authors declare that they have no conflict of interest.
+
+\noindent
+\textbf{Ethical approval} This article does not contain any studies with human participants or animals performed by any of the authors.
 
@@
 
-%\bibliographystyle{abbrvnat}
+\bibliographystyle{sn-basic}
 \bibliography{sn-bibliography}

### 2024-03-21 2024/ITMO paper/sn-article.tex
Add new Results section

@@
 %% Русский язык - закомментировать в финальной версии
-\usepackage[russian]{babel}
+%\usepackage[russian]{babel}
 %% Зачеркивание - закомментировать в финальной версии
-\usepackage[normalem]{ulem}
+%\usepackage[normalem]{ulem}
 
@@
 The reduction is based on the use of space-filling curves (Peano curves).
-These approaches are implemented in the open-source framework of \textit{i}ntelligent \textit{opt}imization methods \textcolor[rgb]{1,0,0}{(iOpt)}.
-To demonstrate the advantages of iOpt, we compare it with the well-known Optuna and HyperOpt frameworks when tuning hyperparameters of various machine learning methods on \textcolor[rgb]{1,0,0}{\sout{a} the}  representative \textcolor[rgb]{1,0,0}{ \sout{set of} } datasets.
+These approaches are implemented in the open-source framework of \textit{i}ntelligent \textit{opt}imization methods (iOpt).
+To demonstrate the advantages of iOpt, we compare it with the well-known Optuna and HyperOpt frameworks when tuning hyperparameters of various machine learning methods on representative datasets.
 The results show that Lipschitz global optimization methods provide comparable (in terms of quality) results in a significantly shorter time compared to known hyperparameter tuning algorithms.}
@@
 
-In this paper, we present a framework of \textit{i}ntelligent \textit{opt}imization methods  \textcolor[rgb]{1,0,0}{iOpt\footnote{\url{https://github.com/aimclub/iOpt/}}}, focused on solving problems of selecting optimal values of hyperparameters of various algorithms (including machine learning \textcolor[rgb]{1,0,0}{and artificial intelligence} methods).
+In this paper, we present a framework of \textit{i}ntelligent \textit{opt}imization methods  iOpt\footnote{\url{https://github.com/aimclub/iOpt/}}, focused on solving problems of selecting optimal values of hyperparameters of various algorithms (including machine learning and artificial intelligence methods).
 
-\textcolor[rgb]{1,0,0}{Отличительной особенностью фреймворка iOpt является использование for hyperparameters tuning методов липшицевой оптимизации, которые являются детерминированными, обладают быстрой сходимостью, а также возможностью адаптации к различного рода специальным постановкам задач.
-Например, рассматриваемые методы допускают обобщение для работы с зашумленными функциями \citep{Sergeyev2020}.
-В данной работе они были адаптированы для решения HPO problems, в которых часть параметров часто являются дискретными и/или категориальными.}
+An essential feature of the iOpt framework is the use of Lipschitz optimisation methods for hyperparameter tuning.
+These methods are deterministic, provide fast convergence, and can be adapted to various types of special problem formulations.
+For example, such methods allow generalisation for working with noisy functions  \citep{Sergeyev2020}.
+In this paper, these methods have been adapted to solve HPO problems where some of the parameters are often discrete and/or categorical.
 
@@
 %Новый текст на русском
-\textcolor[rgb]{1,0,0}{
-Условия сходимости Global Search Algorithm для задач с Categorical Parameters (GSA-CP) могут быть сформулированы в виде следующего утверждения.
+The convergence conditions of the Global Search Algorithm for problems with Categorical Parameters (GSA-CP) can be formulated as the following statement.
 \begin{proposition}
-Пусть функции $ \varphi(y,u_s), y \in D, s = 1,...,S,$ form (\ref{problem_is}) удовлетворяют условию Липшица с соответствующими константами $L_s, s = 1,...,S$, и $L = \max\{L_s\}$.
-И пусть точка $\overline{x}$ be a limit point of the sequence $\left\{x^k\right\} $, полученной при минимизации функции $f(x)$ из (\ref{f}) с помощью алгоритма GSA-CP. Then:
+Let the functions $ \varphi(y,u_s), y \in D, s = 1,...,S,$ from (\ref{problem_is}) satisfy the Lipschitz condition with respective constants $L_s, s = 1,...,S$, and $L = \max\{L_s\}$.
+Also let the point $\overline{x}$ be a limit point of the sequence $\left\{x^k\right\} $, obtained by minimizing the function $f(x)$ from (\ref{f}) using the GSA-CP algorithm. Then:
 \begin{enumerate}
 	\item 
-	если одноверменно с $\overline{x}$ существует другая предельная точка $x'$ последовательности $\left\{x^k\right\}$, тогда $f(\overline{x}) = f(x')$.
+	if at the same time with $\overline{x}$ there exists another limit point $x'$ of the sequence $\left\{x^k\right\}$, then $f(\overline{x}) = f(x')$;
 	\item
-	$z^k=f(x^k) \geq f(\overline{x})$ при любом $k \geq 0$.
+	$z^k=f(x^k) \geq f(\overline{x})$ for any  $k \geq 0$;
 	\item
-	Если на некотором шаге поиска величина $\mu$ from (\ref{mu}) satisfies the condition
+	If at some search step the value $\mu$ from (\ref{mu}) satisfies the condition
 \begin{equation}\label{217}
@@
 \begin{proof}
-Это утверждение доказывается аналогично теореме о сходимости index method of global optimization from \citet{Strongin2000}, принимая во внимание взаимосвясь (\ref{f}) функций  $f(x)$ и $\varphi(y,u_s)$, а также правила (\ref{index}), задающие индексы у точек испытаний.
+This statement is proved in a similar way to the theorem on the convergence of the index method of global optimisation from \citet{Strongin2000}, taking into account the relationship (\ref{f}) of the functions $f(x)$ and $\varphi(y,u_s)$, as well as the rules (\ref{index}), specifying the indices of the trial points.
 \end{proof}
-}
 
@@
 
-\subsection{Results and discussion}
+\subsection{Results on live hyperparameters tuning}
 
@@
 
-% Please add the following required packages to your document preamble:
-% \usepackage{graphicx}
 \begin{table*}
@@
 
-Дополнительная серия экспериментов по сравнению iOpt, Optuna и HyperOpt была проведена с использованием фреймворка JAHS-Bench-201 [ссылка].
-JAHS-Bench предоставляет удобный доступ к набору задач для определения оптимального сочетания гиперпараметров и архитектуры neural network.
-Данный инструмент для сравнения HPO-алгоритмов содержит обширный набор данных о качестве и производительности нейронных сетей, состоящий примерно из 161 million data points, каждая из которых состоит из 20 показателей.
+An additional series of experiments to compare iOpt, Optuna, and HyperOpt were conducted using the JAHS-Bench-201 framework described in \citet{jahsbench}.
+JAHS-Bench provides convenient access to a set of tasks to determine the optimal combination of hyperparameters and neural network architecture.
+This tool for comparing HPO algorithms contains an extensive dataset on the quality and performance of neural networks that comprises approximately 161 million data points, each consisting of 20 performance metrics.
 
@@
 
-Особенностью JAHS-Bench является использование surrogate models fit on performance data, covering different optimization problems (single/multiple objectives), что предоставляет возможность варьировать архитектуру нейросети (выбирая один из имеющихся вариантов), а также ее гиперпараметры.
-При этом использование surrogate models позволяет оценить качество работы обученного классификатора при конкретном наборе гиперпараметров за доли секунды вместо нескольких часов, которые были бы затрачены на обучение реальной нейросети.
-В качестве основы для суррогатных моделей в JAHS-Bench используются три популярных датасета -- CIFAR-10, Colorectal-Histology, Fashion MNIST -- применяемых для задач классификации.
+A particular feature of JAHS-Bench is the use of surrogate models fit on performance data, covering different optimization problems (single/multiple objectives), which makes it possible to modify the neural network architecture (choosing one of the available options) as well as its hyperparameters.
+At the same time, by using surrogate models one can evaluate the performance of a trained classifier for a particular set of hyperparameters in a fraction of a second instead of several hours that would be required to train a real neural network.
+JAHS-Bench uses three popular datasets applied for classification tasks (CIFAR-10, Colorectal-Histology and Fashion MNIST) as the basis for surrogate models.
 
-В проведенных экспериментах настройка гиперпараметров с помощью фреймворков iOpt, Optuna и HyperOpt выполнялась на фиксированной архитектуре нейросети, предложенной в JAHS-Bench качестве архитектуры по умолчанию.
-Для настройки были выбраны традиционно наиболее значимые (с точки зрения исследователя) гиперпараметры: категориальный параметр выбора функции активации (принимающий три возможных значения ReLU, Mish, Hardswish), а также непрерывные параметры Learning Rate and Weight Decay в диапазонах $[10^{-3},1]$ and $[10^{-5},10^{-2}]$ соответственно.
+In our experiments, hyperparameter tuning using iOpt, Optuna and HyperOpt frameworks was performed on the fixed neural network architecture proposed in JAHS-Bench as the default architecture.
+The most relevant (from the researcher's point of view) hyperparameters were chosen for tuning: the categorical parameter of activation function selection (taking on the three possible values ReLU, Mish, Hardswish), as well as the continuous parameters Learning Rate and Weight Decay in the ranges $[10^{-3},1]$ and $[10^{-5},10^{-2}]$ respectively.
 
-В качестве целевой функции использовалась метрика validation accuracy, вычисленная после 200 эпох обучения.
-Точность поиска в критерии остановки всех методов была установлена равной 0.01, а максимальное число испытаний – равным 200.
-Для алгоритмов из Optuna и HyperOpt достигнутые показатели усреднялись по 10 запускам.
-Результаты сравнения представлены в таблице \ref{tab:met-surr}.
-Как и ранее, the values corresponding to the best result on average are highlighted in bold and the values corresponding to the best result for the worst case are highlighted in red.
+The validation accuracy metric calculated after 200 training epochs was used as the objective function.
+The search accuracy in the stopping criterion of all methods was set to 0.01 and the maximum number of trials was set to 200.
+For the algorithms from Optuna and HyperOpt, the obtained data were averaged over 10 runs.
+The comparison results are shown in Table \ref{tab:met-surr}.
+As before, the values corresponding to the best result on average are highlighted in bold and the values corresponding to the best result for the worst case are highlighted in red.
 
-Результаты показывают превосходство iOpt на датасетах CIFAR-10 и Colorectal-Histology.
-При использовании датасета Fashion MNIST фреймворк iOpt превосходит HyperOpt, но уступает Optuna.
-Отметим, что при использовании суррогатных моделей время вычисления целевой функции составляет доли секунды, поэтому результаты, связанные со сравнением времени работы алгоритмов, для этого эксперимента не приводятся.
+The results show the superiority of iOpt on the CIFAR-10 and Colorectal-Histology datasets.
+When using the Fashion MNIST dataset, the iOpt framework outperforms HyperOpt, but is inferior to Optuna.
+Note that when using surrogate models, the computation time of the objective function takes only fractions of a second, so the results related to the comparison of algorithm running times are not reported for this experiment.
 

### 2024-03-29 2024/ITMO paper/sn-article.tex
remove string numbers
submitted paper

@@
 %\documentclass[sn-mathphys-num]{sn-jnl}% Math and Physical Sciences Numbered Reference Style 
-\documentclass[lineno,iicol]{sn-jnl}
+\documentclass[iicol]{sn-jnl}
 %%\documentclass[sn-mathphys-ay]{sn-jnl}% Math and Physical Sciences Author Year Reference Style
@@
 
+\author[2]{\fnm{Lubov} \sur{Yamschikova}}
+
+\author[2]{\fnm{Nikolay O.} \sur{Nikitin}}
+
 \author[1,2]{\fnm{Marina} \sur{Usova}}
@@
 %\orgdiv{Department},
-\orgname{ITMO University}, \orgaddress{\street{Lomonosova St. 9}, \city{St. Petersburg}, \postcode{191002}, 
+\orgname{ITMO University}, \orgaddress{\street{Kronverksky Ave. 49, bldg. A}, \city{St. Petersburg}, \postcode{197101}, 
 %\state{State},
@@
 Currently, complex scientific and/or industrial problems are most often solved by numerical methods.
-Most of such methods, for example, machine learning methods, have hyperparameters, and the choice of specific values of these hyperparameters can determine the efficiency (in a given metric) of the obtained solution under certain constraints on the computation time \citep{Hutter2019,nikitin2021}.
+Most of such methods, for example, machine learning methods, have hyperparameters, and the choice of specific values of these hyperparameters can determine the efficiency (in a given metric) of the obtained solution under certain constraints on the computation time \citep{Hutter2019,nikitin2022automated}.
 The fact that in many cases the difference in the efficiency of the solution obtained with different values of hyperparameters can be quite significant, has resulted in the emergence of the hyperparameter optimization (HPO) problem, as well as the research of approaches to its solution that are implemented in the frameworks for automatic selection of hyperparameters \citep{Tune,optuna,hyperopt,Sherpa}.
@@
 Metaheuristic algorithms only weakly depend on the number of hyperparameters, they use information from previous iterations to perform the current one, but due to the randomness inherent in the methods, they guarantee the finding of the global optimum only in the probabilistic sense.
+
+%Nikitin
+Due to the high complexity of many HPO tasks, the meta-learning techniques are frequently used.
+For example, deep learning models \citep{doke2021survey} are usually too computationally expensive to tune the hyperparameters directly.
+In resent studies, large language models are used as powerful meta-models \citep{zhang2023using}.
+However, the deficiency of meta-learning based approach as not sufficient for many machine learning applications \citep{doke2021survey}.
 
@@
 In particular, the methods have been adapted to solve problems with partially integer parameters, which are very common in HPO problems.
-The experimental results presented in Section \ref{sec_exp} show that the iOpt framework is not inferior in HPO problems to well-known frameworks having the same purpose (such as Optuna \citep{optuna} and HyperOpt \citep{hyperopt}).
+The experimental results presented in Section~\ref{sec_exp} show that the iOpt framework is not inferior in HPO problems to well-known frameworks having the same purpose (such as Optuna~\citep{optuna} and HyperOpt~\citep{hyperopt}).
 
@@
 
-\subsection{Results on live hyperparameters tuning}
+\subsection{Results on ML models}
 
@@
 If one focuses on the total operating time, the performance gain of iOpt is even greater.
-As an example, Figure \ref{fig8} shows a comparison of the hyperparameter tuning times for iOpt, Optuna and Hyperopt on several datasets and different classification models.
+As an example, Fig.~\ref{fig8} shows a comparison of the hyperparameter tuning times for iOpt, Optuna and Hyperopt on several datasets and different classification models.
 For iOpt, the time is shown as a blue dot.
@@
 \centering
-\includegraphics[width=\textwidth]{fig_time}
+\includegraphics[width=0.9\textwidth]{fig_time}
 \caption{Time of hyperparameters tuning.} \label{fig8}
@@
 
-\subsection{Results on surrogate models}
+\subsection{Results on surrogate HPO benchmark}
 
@@
 
+\subsection{Results on AutoML-designed pipelines}
+
+In many cases, HPO is considered not as an independent machine learning task but as a part of automated machine learning workflow (see, e.g., \cite{filippou2023structure}).
+The AutoML task is commonly formulated as CASH -- Combined Algorithms Selection and Hyperparameter Optimization \citep{kim2022survey}.
+Since state-of-the-art AutoML tends to produce complicated ensemble-based modeling pipelines, the efficiency of HPO can differ for AutoML-generated models and static tasks in benchmarks.
+
+In the paper, we choose the FEDOT AutoML framework \citep{nikitin2022automated} as a test bench since it allows the production of diverse pipelines with different complexities.
+The performance of HPO for the tuning of ensemble-based models was compared to the results of optimization algorithms based on Optuna and Hyperopt frameworks.
+Optimizers were compared on ensemble machine-learning pipelines designed by the FEDOT framework.
+In this case, the pipelines are ensembles of ML algorithms assembled together using stacking.
+Such compositions of algorithms can be represented in the form of directed acyclic graphs.
+
+For multi-node pipeline, the optimization strategy can be implemented in different ways:
+\begin{enumerate}
+    \item Optimization of hyperparameters in each node sequentially, maximizing the metric for the parameters of the entire graph.
+    \item Optimization of all hyperparameters of the model simultaneously.
+\end{enumerate}
+
+In this experiment, the second approach was used; the search space was formed from all the hyperparameters of each of the vertices of the graph.
+
+The algorithms were compared for regression task. %two machine learning tasks: regression and classification. 
+The optimizers were run for a limited number of iterations (20 and 100).
+Each of the experiments was conducted with a different set of test models, datasets, and algorithms and was run for 30 times.
+Based on the results of the experiments, the average increase in the target metric on test data was compared, that is, the difference in metrics for models with the default hyperparameters and with the hyperparameters obtained after optimization.
+
+%\subsubsection{Regression}
+For the regression problem, three directed cyclic graphs were chosen to represent the different pipeline structures for the regression problem.
+These graphs differed in the complexity of their structure and the number of hyperparameters for optimization.
+The following types of graphs were taken as test ones: linear, branched, and complex (see Fig.~\ref{fig_reg_pipe}).
+
+\begin{figure}
+\center{\includegraphics[width=1.0\linewidth]{regression_pipe.png} }
+\caption{Test graphs for regression task: a) linear pipeline, where \textit{knnreg} -- k-nearest neighbors algorithm, \textit{ridge} -- linear regression with L2 regularization; b) branched pipeline, where \textit{svr} -- support vector machine, \textit{rfr} -- random forest algorithm; c) complex pipeline, where \textit{lasso} -- linear regression with L1 regularization.}\label{fig_reg_pipe}
+\end{figure}
+
+Optimization of the hyperparameters for each pipeline was conducted for three data sets:
+\begin{enumerate}
+    \item California Housing Prices (CHP) -- data contains information from the 1990 California Census and provides average home prices for various quarters.
+The average cost of houses is the target variable.
+    \item Delta Ailerons (DA) -- this data set is derived from the F16 aircraft aileron control problem.
+The attributes describe the state of the aircraft, and the goal is to predict the control input on the aircraft's ailerons.
+    \item Pol -- a data set with forty-eight features describing a telecommunications problem.
+\end{enumerate}
+
+\begin{table*}
+\centering
+\caption{Metric improvement for regression task, where L -- linear pipeline, B -- branched pipeline, C -- complex pipeline.}
+\label{tab:automl_reg}
+\begin{tabular}{llccccc} \hline
+Dataset & № iters & Test graph& iOpt & Hyperopt(tpe) & Optuna(tpe) & p-value\\ \hline
+\multirow{6}{*}{CHP} & \multirow{3}{*}{20} & L& 1.2 & 1.2$\pm$0.1& 0.9$\pm$0.6& 0.9048 \\
+ &  & B& 2.2& 2.2$\pm$0.3& \textbf{2.5$\pm$0.3}& 0.0001\\
+ &  & C& \textbf{4.6} & 3.2$\pm$0.4& 3.2$\pm$0.6& 0.0000 \\
+ & \multirow{3}{*}{100} & L
+& 1.1& 1.1$\pm$0.0& 1.1$\pm$0.2& -\\
+ &  & B
+& 2.0& 2.1$\pm$0.0& \textbf{2.5$\pm$0.3}& 0.0000\\
+ &  & C& \textbf{4.6} & 4.1$\pm$0.3& 4.0$\pm$0.6& 0.0000 \\ \hline
+\multirow{6}{*}{DA} & \multirow{3}{*}{20} & L 
+& \textbf{4.2} & 0.5$\pm$1.8& 0.5$\pm$1.1& 0.0000 \\ 
+ &  & B
+& \textbf{1.7}& 1.4$\pm$0.1& 1.4$\pm$0.1& 0.0000\\
+ &  & C& \textbf{62.3} & 33.5$\pm$31.4& 11.9$\pm$23.6& 0.0001 \\
+ & \multirow{3}{*}{100} & L
+& 4.2 & \textbf{6.3$\pm$2.1}& 6.1$\pm$2.3& 0.0005 \\
+ &  & 
+B& \textbf{1.7}& 1.5$\pm$0.2& 1.5$\pm$0.2& 0.0000\\
+ &  & C& 62.3 & \textbf{63.3$\pm$1.4}& 50.6$\pm$24.3& 0.0065 \\ \hline
+\multirow{6}{*}{Pol} & \multirow{3}{*}{20} & 
+L
+& \textbf{1.2} & 0.8$\pm$0.3& 0.4$\pm$0.6& 0.0000 \\
+ &  & 
+B& \textbf{19.0}& 1.5$\pm$2.2& 3.7$\pm$0.35.4& 0.0006\\
+ &  & C& 0.0 & 0.0$\pm$0.0& 0.0$\pm$0.0& -  \\ 
+ & \multirow{3}{*}{100} & L& \textbf{1.2} & \textbf{1.2$\pm$0.0}& 1.0$\pm$0.5& 0.0000 \\
+ &  &  B& \textbf{19.0}& 17.8$\pm$2.3& 13.9$\pm$5.3& 0.0000\\
+ &  & C& 0.0 & 0.0$\pm$0.0& 0.0$\pm$0.0& - \\ \hline
+\end{tabular}
+\end{table*}
+
+The results and their statistical significance (Friedman and Nemenyi tests) are shown in Table~\ref{tab:automl_reg}.
+The experiments confirmed that the performance of iOpt in the HPO problem for ensemble models is comparable to Optuna, however, the undisputed advantage of the iOpt algorithm is its deterministic nature.
+
 \section{Conclusion}\label{sec13}

### 2024-06-17 2024/OPTIMA/OPTIMA_2024_MCO.tex
Optima MCO

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{cite}
\usepackage{hyperref}

\usepackage[misc,geometry]{ifsym}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Sergey Konnov\orcidID{0009-0003-4590-0870} \and
Evgeny Kozinov \Letter\orcidID{0000-0001-6776-0096} \and
Konstantin Barkalov \Letter\orcidID{0000-0001-5273-2471} \and
Vladimir Grishagin\orcidID{0000-0002-2884-3670}}
%
\authorrunning{S. Konnov, E. Kozinov, K. Barkalov and V. Grishagin}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.

\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
\email{a230172@unn.ru}, \email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, \email{vagris@unn.ru}}
%
\maketitle

%
\begin{abstract}
To solve a multicriterial optimization problem, it is necessary to find a whole set of parameters values corresponding to non-dominated criteria values (Pa-reto set). The complexity of these problems increases significantly in the case of the criteria are multiextremal. In the paper discusses several varia-tions of the multicriteria optimization algorithm using the black box method with multiextremal criteria. The developed algorithms are based on an infor-mation-statistical approach and a machine learning procedure used to in-crease the efficiency of constructing the Pareto set. As a novelty, the authors proposed a new approach to using the constructed machine learning models. The method involves using the calculated probability of a trial points belong-ing to the Pareto set to assess the prospects one. The effectiveness of differ-ent variations of the developed algorithms estimated in a representative com-putational experiment.

\keywords{Multicriterial Problems \and Global Optimization \and Machine Learning.}
\end{abstract}
%
%
%
\section{Introduction}
\label{sec:1}

\cite{Miettinen1999,Ehrgott2005,Pardalos2017,ML_MCO_2023,Evtushenko2014,Deb2002,Durillo2010,Mostaghim2007,NDG09,RC05,ZLT01,Gergel2019_2,Gergel2018,GergelKozinov2020,Marler2004,Strongin2000,Sergeyev2013,SVM_2000,PROB_2004,iOpt_url,Grishagin2015_2}

\section{Problem statement}
\label{sec:2}

\begin{equation}
\label{eq:01}
  \Phi(y) = (f_1 (y),f_2 (y), \dots, f(y)) \to \min, y \in D
\end{equation}

\begin{equation}
\label{eq:02}
    D=\{y \in R^n : a_i \leq y_i \leq b_i, 1 \leq i \leq N\}
\end{equation}

\begin{equation}
\label{eq:03}
|f_i (y') - f_i (y'')| \leq L_i \|y' - y''\| ,y',y'' \in D, 1 \leq i \leq s,
\end{equation}

\section{General computational scheme}
\label{sec:3}

\begin{equation}
\label{eq:04}
\min \varphi(y) = F( \lambda, y ), y \in D,
\end{equation}

\begin{equation}
\label{eq:05}
F(\lambda, y) = \max_{1 \leq i \leq s} {(\lambda_i f_i (y))}, y \in D, \sum_{i=1}^s {\lambda_i} = 1, 0 \leq \lambda_i \leq 1.
\end{equation}

\begin{equation}
\label{eq:06}
F(y) = \{ F(\lambda_1,y),F(\lambda_2,y), \dots ,F(\lambda_\tau,y)\}
\end{equation}

\begin{equation}
\label{eq:07}
\min{\varphi(y(x))} = \min {\varphi(y)}, x \in [0,1], y \in D
\end{equation}

\begin{equation}
\label{eq:08}
|\varphi(y(x')) - \varphi(y(x''))| \leq H \|x' - x''\|^{1/N} , x \in [0,1],
\end{equation}

\begin{equation}
    \label{eq:09}
    x_0 = 0 < x_1 < \dots < x_i < \dots < x_{k} = 1.
\end{equation}

\begin{equation}
    \label{eq:10}
		\begin{matrix}
		m=\begin{cases}
				\begin{matrix}
					 r M, & M >0 \\
					 1, & M = 0 
				\end{matrix} 
			\end{cases} ,
		M = \max_{1 \leq i \leq k} \frac{| z_i - z_{i-1}|}{\varrho_i}, \\
		z_i = \varphi( x_i ), \varrho_i=\sqrt[N]{x_i-x_{i-1}}
		\end{matrix}
\end{equation}

\begin{equation}
    \label{eq:11}
    R(j) = \varrho_j + \frac{(z_i-z_{i-1})^2}{m^2 \varrho_i} - \frac{2 (z_i+z_{i-1})}{m}, 1 \leq i \leq k.
\end{equation}

\begin{equation}
    \label{eq:12}
    R(t) = argmax_{1 \leq i \leq k} {R(i)}.
\end{equation}

\begin{equation}
    \label{eq:13}
    x^{k+1} = \frac{x_t + x_{t-1}}{2} - sign(z_t - z_{t-1}) \frac{1}{2r} \left(\frac{|z_l - z_{l-1}|}{m} \right)^N,
\end{equation}

\begin{equation}
    \label{eq:14}
    \varrho < \varepsilon.
\end{equation}

\section{Approaches to improving search efficiency}
\label{sec:4}

\begin{equation}
    \label{eq:15}
    \Omega_k=\{\mu^i = (y^i,f^i=f(y^i))^T: 0 \leq i \leq k\}.
\end{equation}

\begin{equation}
    \label{eq:16}
    R(i) = R_{gsa} (i) +  \alpha R_{ps} (i).
\end{equation}

\begin{equation}
    \label{eq:17}
d'_i=
\begin{cases}
  \begin{matrix}
     d_i / d_{max}, & d_i > 0 \\
     -d_i / d_{min}, & d_i < 0 
  \end{matrix}
\end{cases}, 
1 \leq i \leq k,
\end{equation}

\begin{equation}
    \label{eq:18}
p'_i= \frac{ \log (p_i) - \log (p_{min})}{ \log (p_{max}) - \log (p_{min})} , 1 \leq i \leq k,
\end{equation}

\section{Results of computational experiments}
\label{sec:5}

\begin{equation}
    \label{eq:19}
		\begin{matrix}
		  f(y)= -(AB + CD)^{1/2} \\
			AB =(\sum_{i=1}^7{\sum_{j=1}^7{[A_{ij} a_{ij} (y_1,y_2) + B_{ij} b_{ij} (y_1,y_2)])}})^2 \\
			CD =(\sum_{i=1}^7{\sum_{j=1}^7{[C_{ij} a_{ij} (y_1,y_2) - D_{ij} b_{ij} (y_1,y_2)])}})^2 \\
			a_{ij} (y_1,y_2) = \sin(\pi i y_1) \sin(\pi j y_2), \\
			b_{ij} (y_1,y_2) = \cos(\pi i y_1) \cos(\pi j y_2),
		\end{matrix}
\end{equation}

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
 \bibliographystyle{splncs04}
 \bibliography{bibliography}
%
%\begin{thebibliography}{8}
%\bibitem{ref_article1}
%Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
%
%\bibitem{ref_lncs1}
%Author, F., Author, S.: Title of a proceedings paper. In: Editor,
%F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
%Springer, Heidelberg (2016). \doi{10.10007/1234567890}
%
%\bibitem{ref_book1}
%Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
%Location (1999)
%
%\bibitem{ref_proc1}
%Author, A.-B.: Contribution title. In: 9th International Proceedings
%on Proceedings, pp. 1--2. Publisher, Location (2010)
%
%\bibitem{ref_url1}
%LNCS Homepage, \url{http://www.springer.com/lncs}, last accessed 2023/10/25
%\end{thebibliography}
\end{document}

### 2024-06-17 2024/OPTIMA/OPTIMA_2024_MCO.tex
add titlerunning
fix references list

@@
 %
-\title{Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems}
+\title{Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.
+FSWR-2023-0034, and by by the Research and Education Mathematics Centre ``Mathematics for Future Technologies'', project no.
+075-02-2024-1439.} }
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Algorithms for solving multicriterial problems}
 % If the paper title is too long for the running head, you can set
@@
 \begin{abstract}
-To solve a multicriterial optimization problem, it is necessary to find a whole set of parameters values corresponding to non-dominated criteria values (Pa-reto set).
+To solve a multicriterial optimization problem, it is necessary to find a whole set of parameters values corresponding to non-dominated criteria values (Pareto set).
 The complexity of these problems increases significantly in the case of the criteria are multiextremal.
 In the paper discusses several varia-tions of the multicriteria optimization algorithm using the black box method with multiextremal criteria.
-The developed algorithms are based on an infor-mation-statistical approach and a machine learning procedure used to in-crease the efficiency of constructing the Pareto set.
+The developed algorithms are based on an information-statistical approach and a machine learning procedure used to increase the efficiency of constructing the Pareto set.
 As a novelty, the authors proposed a new approach to using the constructed machine learning models.
-The method involves using the calculated probability of a trial points belong-ing to the Pareto set to assess the prospects one.
-The effectiveness of differ-ent variations of the developed algorithms estimated in a representative com-putational experiment.
+The method involves using the calculated probability of a trial points belonging to the Pareto set to assess the prospects one.
+The effectiveness of different variations of the developed algorithms estimated in a representative computational experiment.
 

### 2024-06-25 2024/OPTIMA/OPTIMA_2024_MCO.tex
fig added
Optima part 1

@@
 \begin{abstract}
-To solve a multicriterial optimization problem, it is necessary to find a whole set of parameters values corresponding to non-dominated criteria values (Pareto set).
-The complexity of these problems increases significantly in the case of the criteria are multiextremal.
-In the paper discusses several varia-tions of the multicriteria optimization algorithm using the black box method with multiextremal criteria.
-The developed algorithms are based on an information-statistical approach and a machine learning procedure used to increase the efficiency of constructing the Pareto set.
-As a novelty, the authors proposed a new approach to using the constructed machine learning models.
-The method involves using the calculated probability of a trial points belonging to the Pareto set to assess the prospects one.
-The effectiveness of different variations of the developed algorithms estimated in a representative computational experiment.
+To solve a multicriterial optimization problem, it is necessary to find a whole set of independent variables corresponding to non-dominated criteria values (Pareto set).
+The complexity of these problems increases significantly in the case when the criteria are multiextremal.
+The paper presents several variations of the global optimization algorithm intended for solving black-box multiobjective problems with multiextremal criteria.
+The developed methods are based on a combination of the information-statistical approach and machine learning procedures aimed at increasing the efficiency of constructing the Pareto set.
+As a novelty, the paper describes a new approach to using machine learning models within the algorithm of multicriterial optimization.
+The effectiveness of different versions of the developed algorithm is estimated on the base of  a representative computational experiment.
 
@@
 \cite{Miettinen1999,Ehrgott2005,Pardalos2017,ML_MCO_2023,Evtushenko2014,Deb2002,Durillo2010,Mostaghim2007,NDG09,RC05,ZLT01,Gergel2019_2,Gergel2018,GergelKozinov2020,Marler2004,Strongin2000,Sergeyev2013,SVM_2000,PROB_2004,iOpt_url,Grishagin2015_2}
+
+Problems of optimal choice are widespread in many areas of activity.
+As a rule, in complicated decision making models there are several functions to be optimized which reflect the decision making objectives and those are, in general case, contradictory.
+In these circumstances, the notion of optimal solution is more complicated than in single criterion (scalar) statement formulated as mathematical programming problem because there exist incomparable decisions.
+In this situation under introduced definition of domination, the set of non-dominated variants (Pareto set) is taken as the solution of the multiobjective problem.
+
+In our study we deal with methods for solving the most complex class of multicriterial problems where criteria are considered as ``black-box'' functions and are multiextremal \cite{Miettinen1999,Ehrgott2005,Pardalos2017,Strongin2000,Sergeyev2013}.
+
+It should be noted that problems of multiextremal optimization are among the most complicated problems of mathematical programming.
+Unlike the local optimization where an optimal point has to be juxtaposed with other points only in its region of attraction, the global solution must be better w.r.t.
+all points in the feasible domain.
+If to solve such problems by grid search methods, the amount of computations will rise exponentially when increasing the number of parameters (variables of the problem).
+
+For solving the multiextremal problems the more effective algorithms are used which can be divided into two key classes, namely, into deterministic and meta-heuristic (nature-inspired) methods \cite{ML_MCO_2023,Evtushenko2014,Deb2002,Durillo2010,Mostaghim2007,NDG09,RC05,ZLT01,Gergel2018,GergelKozinov2020}.
+
+Meta-heuristic algorithms model the behavior of living organisms or nature processes \cite{Deb2002,Durillo2010,Mostaghim2007,NDG09,RC05,ZLT01}.
+As a rule, these methods are simple enough in implementation and can demonstrate good efficiency for unimodal objective functions and multiextremal problems with a small number of local optima; however, they do not ensure finding the best solution.
+
+Deterministic algorithms guarantee convergence to global optimum, but their computational schemes are more complex.
+However, advanced deterministic algorithms outperform metaheuristic ones in terms of quality \cite{Kvasov2018,Sergeyev2018}.
+
+One of the qualitative deterministic methods for solving multiextremal problems is the  information-statistical algorithm of global search \cite{Strongin2000,Sergeyev2013}.
+This algorithm proposed initially for solving scalar problems was successfully developed for analyzing multiobjective statements \cite{ML_MCO_2023,Gergel2018,GergelKozinov2020}.
+
+In the publication \cite{ML_MCO_2023} we proposed an approach to solving multicriterial optimization problems combining the classic information-statistical global search algorithm with machine learning methods.
+In the framework of the approach a problem of multicriterial optimization was transformed to a series of scalar optimization subproblems.
+In the course of solving scalar subproblems, on the base of machine learning methods a hyperplane was built and gradually refined which separated points of effective and ineffective solutions in the criteria space.
+Additional information connected with the separating hyperplane was embedded in the computational scheme of the optimization algorithm and, as a result, the number of trials (criteria evaluations) in domains with ineffective solutions was significantly decreased.
+The experimental results showed that in comparison with a number of heuristic algorithms, the search for effective solutions was several times faster.
+At the same time, this approach can be improved because utilizing only the separating plane limits the class of machine learning techniques which can be applied for enhancing the global search algorithm.
+
+In the framework of our study we propose a new approach to separating search trial points in the criteria space based on the probability that reflects the belonging of the trial point to the Pareto set.
+The most machine learning models allow obtaining this metric which makes it possible to choose an algorithm for building a machine learning model depending on the features of the problem being solved.
+
+The paper has the following structure.
+Section \ref{sec:2} describes the mathematical statement of the problem to be studied.
+Section \ref{sec:3} is devoted to a brief description of the basic algorithms used for solving problems of multicriterial optimization.
+Section \ref{sec:4} contains information on the techniques applied for reducing the number of trials during the optimization process.
+Subsection \ref{ssec:41} gives a short description of the method considered in \cite{ML_MCO_2023}.
+A new approach to using probabilities of trial points belonging to different efficiency classes is presented in subsection \ref{ssec:42}.
+Section \ref{sec:5} contains results of computational experiments demonstrating the effectiveness of the proposed approach.
 
@@
 
+\subsection{Section 41}
+\label{ssec:41}
+
+\subsection{Section 42}
+\label{ssec:42}
+
 \begin{equation}
@@
 
+\begin{figure}
+\center
+\includegraphics[width=0.8\textwidth]{fig1.png}
+\label{fig1}
+\caption{A figure caption.} 
+\end{figure}
+
 \begin{equation}
@@
 \end{equation}
+
+\begin{figure}
+\center
+\includegraphics[width=0.8\textwidth]{fig2.png}
+\label{fig2}
+\caption{A figure caption.} 
+\end{figure}
 
@@
 \end{equation}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{fig3.png}
+\label{fig3}
+\caption{A figure caption.} 
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{fig4.png}
+\label{fig4}
+\caption{A figure caption.} 
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{fig5.png}
+\label{fig5}
+\caption{A figure caption.} 
+\end{figure}
 

### 2024-06-25 2024/OPTIMA/OPTIMA_2024_MCO.tex
add some references

@@
 
-For solving the multiextremal problems the more effective algorithms are used which can be divided into two key classes, namely, into deterministic and meta-heuristic (nature-inspired) methods \cite{ML_MCO_2023,Evtushenko2014,Deb2002,Durillo2010,Mostaghim2007,NDG09,RC05,ZLT01,Gergel2018,GergelKozinov2020}.
+For solving the multiextremal problems the more effective algorithms are used which can be divided into two key classes, namely, into deterministic (see, e.g., \cite{Evtushenko2014,Gergel2018,GergelKozinov2020,Paulavicius2020,Jones2021}) and meta-heuristic (see, e.g., \cite{Battiti2009,Gendreau2010,Eiben2015}) methods .
 
@@
 
-In the publication \cite{ML_MCO_2023} we proposed an approach to solving multicriterial optimization problems combining the classic information-statistical global search algorithm with machine learning methods.
+In the paper \cite{ML_MCO_2023} we proposed an approach to solving multicriterial optimization problems combining the classic information-statistical global search algorithm with machine learning methods.
 In the framework of the approach a problem of multicriterial optimization was transformed to a series of scalar optimization subproblems.

### 2024-06-26 2024/OPTIMA/OPTIMA_2024_MCO.tex
OPTIMA part 2

@@
 
+The multicriterial optimization (MCO) problem can be formulated as the minimization problem of the vector-function. Let us introduce the vector-function (vector criterion)
+
 \begin{equation}
 \label{eq:01}
-  \Phi(y) = (f_1 (y),f_2 (y), \dots, f(y)) \to \min, y \in D
+  \Phi(y) = (f_1 (y),f_2 (y), \dots, f_s(y))
+\end{equation}
+where scalar functions $f_i(y)$, $1 \leq i \leq s$, are considered as the partial efficiency criteria, $y=(y_1,y_2, \dots ,y_N)$ is the vector of independent parameters belonging to the feasible domain $D \in R^N$ that is the hyperparallelepiped
+\begin{equation}
+\label{eq:02}
+    D=\{y \in R^N : a_i \leq y_i \leq b_i, 1 \leq i \leq N\}
+\end{equation}
+in the $N$-dimensional Euclidean space, where $a_i$, $b_i$, $1 \leq i \leq N$, are constant values.
+
+The best value of the partial criterion $f_i(y)$, $1 \leq i \leq s$, is supposed to correspond to the global minimum of this criterion in the domain $D$.
+
+We will denote this problem in the form
+\begin{equation}
+\label{eq:02a}
+  \Phi(y) = (f_1 (y),f_2 (y), \dots, f_s(y)) \to \min, y \in D
 \end{equation}
 
-\begin{equation}
-\label{eq:02}
-    D=\{y \in R^n : a_i \leq y_i \leq b_i, 1 \leq i \leq N\}
-\end{equation}
+In the framework of this study, one of the most complicated classes of the above statement is considered in which the criteria $f_i(y)$, $1 \leq i \leq s$, are multiextremal and presented as the ``black-box'' functions.
+Moreover, these criteria are supposed to satisfy the Lipschitz condition
 
@@
 \end{equation}
+where Lipschitz constants $L_i>0$, $1 \leq i \leq s$, are unknown \textit{a priopi}.
+
+Without loss of generality, the values of partial criteria will be supposed to be positive.
+If a partial criterion does not satisfy this requirement, it is easy to build a new criterion by means of adding to it a sufficiently large positive constant so that replacing in the problem (\ref{eq:02a}) the initial criterion with the modified one will not change its solution.
+Note that such constant can be selected in the course of optimization \cite{ML_MCO_2023}.
+
+As it was mentioned above, the partial criteria contradictoriness leads to the necessity to introduce the set of non-dominated points   called Pareto set as the general solution of the multicriterial problem \cite{Miettinen1999,Ehrgott2005,Pardalos2017}.
+Hereinafter, any point of the Pareto set will be called a partial solution or just an effective point.
 

### 2024-06-26 2024/OPTIMA/OPTIMA_2024_MCO.tex
fix some misprints

@@
 %
-\title{Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.
+\title{Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriteria problems\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.
 FSWR-2023-0034, and by by the Research and Education Mathematics Centre ``Mathematics for Future Technologies'', project no.
@@
 %
-\titlerunning{Algorithms for solving multicriterial problems}
+\titlerunning{Algorithms for solving multicriteria problems}
 % If the paper title is too long for the running head, you can set
@@
 \begin{abstract}
-To solve a multicriterial optimization problem, it is necessary to find a whole set of independent variables corresponding to non-dominated criteria values (Pareto set).
+To solve a multicriteria optimization problem, it is necessary to find a whole set of independent variables corresponding to non-dominated criteria values (Pareto set).
 The complexity of these problems increases significantly in the case when the criteria are multiextremal.
@@
 The developed methods are based on a combination of the information-statistical approach and machine learning procedures aimed at increasing the efficiency of constructing the Pareto set.
-As a novelty, the paper describes a new approach to using machine learning models within the algorithm of multicriterial optimization.
+As a novelty, the paper describes a new approach to using machine learning models within the algorithm of multicriteria optimization.
 The effectiveness of different versions of the developed algorithm is estimated on the base of  a representative computational experiment.
 
-\keywords{Multicriterial Problems \and Global Optimization \and Machine Learning.}
+\keywords{Multicriteria Problems \and Global Optimization \and Machine Learning.}
 \end{abstract}
@@
 
-In our study we deal with methods for solving the most complex class of multicriterial problems where criteria are considered as ``black-box'' functions and are multiextremal \cite{Miettinen1999,Ehrgott2005,Pardalos2017,Strongin2000,Sergeyev2013}.
+In our study we deal with methods for solving the most complex class of multicriteria problems where criteria are considered as ``black-box'' functions and are multiextremal \cite{Miettinen1999,Ehrgott2005,Pardalos2017,Strongin2000,Sergeyev2013}.
 
@@
 
-In the paper \cite{ML_MCO_2023} we proposed an approach to solving multicriterial optimization problems combining the classic information-statistical global search algorithm with machine learning methods.
-In the framework of the approach a problem of multicriterial optimization was transformed to a series of scalar optimization subproblems.
+In the paper \cite{ML_MCO_2023} we proposed an approach to solving multicriteria optimization problems combining the classic information-statistical global search algorithm with machine learning methods.
+In the framework of the approach a problem of multicriteria optimization was transformed to a series of scalar optimization subproblems.
 In the course of solving scalar subproblems, on the base of machine learning methods a hyperplane was built and gradually refined which separated points of effective and ineffective solutions in the criteria space.
@@
 Section \ref{sec:2} describes the mathematical statement of the problem to be studied.
-Section \ref{sec:3} is devoted to a brief description of the basic algorithms used for solving problems of multicriterial optimization.
+Section \ref{sec:3} is devoted to a brief description of the basic algorithms used for solving problems of multicriteria optimization.
 Section \ref{sec:4} contains information on the techniques applied for reducing the number of trials during the optimization process.
@@
 
-The multicriterial optimization (MCO) problem can be formulated as the minimization problem of the vector-function. Let us introduce the vector-function (vector criterion)
+The multicriteria optimization (MCO) problem can be formulated as the minimization problem of the vector-function. Let us introduce the vector-function (vector criterion)
 
@@
 \end{equation}
-in the $N$-dimensional Euclidean space, where $a_i$, $b_i$, $1 \leq i \leq N$, are constant values.
-
+in the $N$-dimensional Euclidean space, and $a_i$, $b_i$, $1 \leq i \leq N$, are constant values.
 The best value of the partial criterion $f_i(y)$, $1 \leq i \leq s$, is supposed to correspond to the global minimum of this criterion in the domain $D$.
@@
 
-As it was mentioned above, the partial criteria contradictoriness leads to the necessity to introduce the set of non-dominated points   called Pareto set as the general solution of the multicriterial problem \cite{Miettinen1999,Ehrgott2005,Pardalos2017}.
-Hereinafter, any point of the Pareto set will be called a partial solution or just an effective point.
+As it was mentioned above, the partial criteria contradictoriness leads to the necessity to introduce the set of non-dominated points called Pareto set as the general solution of the multicriteria problem \cite{Miettinen1999,Ehrgott2005,Pardalos2017}.
+Hereinafter, any point of the Pareto set will be called a \textit{partial solution} or just an \textit{efficient point}.
 

### 2024-06-27 2024/OPTIMA/OPTIMA_2024_MCO.tex
Optima part 3

@@
 
+For solving the optimization problem (\ref{eq:02a}), in our study the computational scheme \cite{ML_MCO_2023} has been applied. It consists of the following main stages.
+
+\begin{enumerate}
+	\item Scalarization of the vector criterion (\ref{eq:01}).
+	\item Dimensionality reduction.
+	\item Optimization of transformed functions.
+\end{enumerate}
+Below is a brief description of each stage.
+
+\textbf{Scalarization.} It is possible to find the Pareto set of optimal parameters either by solving the problem directly in the initial statement \cite{Evtushenko2014,Deb2002,Durillo2010,Mostaghim2007,NDG09,RC05,ZLT01} or by transforming it to a series of scalar subproblems when each subproblem provides finding separate efficient point \cite{Pardalos2017,ML_MCO_2023,Gergel2019_2,Gergel2018,GergelKozinov2020}.
+
+Within the framework of the described research, the second approach is used.
+In general, the optimization problem generated by scalarization of the vector criterion (\ref{eq:01}) can be written in the form
+
 \begin{equation}
@@
 \end{equation}
+where $F$ is a scalar function including partial criteria $f_i$, $1 \leq i \leq s$, and the vector
+\begin{equation}
+\label{eq:04a}
+\lambda = (\lambda_1, \lambda_2, \dots, \lambda_s), \lambda_i > 0, 1 \leq i \leq s, \sum_{i = 1}^s {\lambda_i} = 1,
+\end{equation}
+reflects the ``importance'' (weights) of corresponding partial criteria.
 
+There are various ways to construct the function $F$ \cite{Miettinen1999,Ehrgott2005,GergelKozinov2020,Marler2004}. In our study we utilized the maximum convolution 
 \begin{equation}
@@
 
-\begin{equation}
-\label{eq:06}
-F(y) = \{ F(\lambda_1,y),F(\lambda_2,y), \dots ,F(\lambda_\tau,y)\}
-\end{equation}
+%\begin{equation}
+%\label{eq:06}
+%F(y) = \{ F(\lambda_1,y),F(\lambda_2,y), \dots ,F(\lambda_\tau,y)\}
+%\end{equation}
 
+If all the partial criteria are Lipschitzian and at least one function $f_i(y)$, $1 \leq i \leq s$, is multiextremal, the convolution (\ref{eq:05}) will be Lipschitzian and multiextremal as well.
+
+In the case of partial criteria positivity the global minimizer of the convolution (\ref{eq:05}) is the Pareto-optimal point.
+In order to obtain all elements of the Pareto set, it is necessary to solve the series of scalar problems (\ref{eq:04}) for all weight vectors $\lambda$ satisfying the conditions from (\ref{eq:05}).
+
+\textbf{Dimensionality reduction.} For solving the optimization problems (\ref{eq:05}) the dimensionality reduction \cite{Gergel2019_2,Gergel2018,GergelKozinov2020} on the base of Peano curves is applied.
+The Peano curve $y(x)$ maps the close interval $[0,1]$ of the real axis onto the parallelepiped (\ref{eq:02}) unambiguously and continuously.
+It enables to replace the multidimensional problem (\ref{eq:04}) with solving the equivalent univariate problem
 \begin{equation}
@@
 
+It should be noted that if the convolution $F(\lambda,y)$ satisfies the Lipschitz condition then the reduced function (\ref{eq:02}) meets the H{\" o}lder condition
+
 \begin{equation}
@@
 \end{equation}
+with the H{\" o}lder constant $H=2L\sqrt{N+3}$, where $L$ is the Lipschitz constant of $F(\lambda,y)$ in the domain $D$.
 

### 2024-06-27 2024/OPTIMA/OPTIMA_2024_MCO.tex
fix some misprints

@@
 For solving the optimization problem (\ref{eq:02a}), in our study the computational scheme \cite{ML_MCO_2023} has been applied. It consists of the following main stages.
-
  \begin{enumerate}
@@
 
-\textbf{Scalarization.} It is possible to find the Pareto set of optimal parameters either by solving the problem directly in the initial statement \cite{Evtushenko2014,Deb2002,Durillo2010,Mostaghim2007,NDG09,RC05,ZLT01} or by transforming it to a series of scalar subproblems when each subproblem provides finding separate efficient point \cite{Pardalos2017,ML_MCO_2023,Gergel2019_2,Gergel2018,GergelKozinov2020}.
+\textbf{Scalarization.} It is possible to find the Pareto set either by solving the problem directly in the initial statement \cite{Evtushenko2014,Deb2002,Durillo2010,Mostaghim2007,NDG09,RC05,ZLT01} or by transforming it to a series of scalar subproblems when each subproblem provides finding separate efficient point \cite{Pardalos2017,ML_MCO_2023,Gergel2019_2,Gergel2018,GergelKozinov2020}.
 
@@
 In general, the optimization problem generated by scalarization of the vector criterion (\ref{eq:01}) can be written in the form
-
 \begin{equation}
 \label{eq:04}
-\min \varphi(y) = F( \lambda, y ), y \in D,
+\min \left\{\varphi(y) = F( \lambda, y ), y \in D\right\},
 \end{equation}
@@
 
-There are various ways to construct the function $F$ \cite{Miettinen1999,Ehrgott2005,GergelKozinov2020,Marler2004}. In our study we utilized the maximum convolution 
+There are various ways to construct the function $F(\lambda, y)$ \cite{Miettinen1999,Ehrgott2005,GergelKozinov2020,Marler2004}. In our study we utilized the maximum convolution 
 \begin{equation}
 \label{eq:05}
-F(\lambda, y) = \max_{1 \leq i \leq s} {(\lambda_i f_i (y))}, y \in D, \sum_{i=1}^s {\lambda_i} = 1, 0 \leq \lambda_i \leq 1.
+F(\lambda, y) = \max_{1 \leq i \leq s} \left\{\lambda_i f_i (y)\right\}, y \in D, \sum_{i=1}^s {\lambda_i} = 1, 0 \leq \lambda_i \leq 1.
 \end{equation}
@@
 \label{eq:07}
-\min{\varphi(y(x))} = \min {\varphi(y)}, x \in [0,1], y \in D
+\min \left\{\varphi(y), y \in D\right\} = \min\left\{ \varphi(y(x)), x \in [0,1] \right\}.
 \end{equation}
 
-It should be noted that if the convolution $F(\lambda,y)$ satisfies the Lipschitz condition then the reduced function (\ref{eq:02}) meets the H{\" o}lder condition
-
+It should be noted that if the convolution $F(\lambda,y)$ satisfies the Lipschitz condition then the reduced function $\varphi(y(x)) = F(\lambda,y(x))$ meets the H{\" o}lder condition
 \begin{equation}

### 2024-06-28 2024/OPTIMA/OPTIMA_2024_MCO.tex
OPTIMA part 4

@@
 
-If all the partial criteria are Lipschitzian and at least one function $f_i(y)$, $1 \leq i \leq s$, is multiextremal, the convolution (\ref{eq:05}) will be Lipschitzian and multiextremal as well.
-
 In the case of partial criteria positivity the global minimizer of the convolution (\ref{eq:05}) is the Pareto-optimal point.
@@
 
-\textbf{Dimensionality reduction.} For solving the optimization problems (\ref{eq:05}) the dimensionality reduction \cite{Gergel2019_2,Gergel2018,GergelKozinov2020} on the base of Peano curves is applied.
+If all the partial criteria are Lipschitzian and at least one function $f_i(y)$, $1 \leq i \leq s$, is multiextremal, the convolution (\ref{eq:05}) will be Lipschitzian and multiextremal as well.
+Thus, global search algorithms must be used to solve problems (\ref{eq:04}) with convolution (\ref{eq:05}).
+
+\textbf{Dimensionality reduction.} For solving the optimization problems (\ref{eq:04}) the dimensionality reduction \cite{Gergel2019_2,Gergel2018,GergelKozinov2020} on the base of Peano curves is applied.
 The Peano curve $y(x)$ maps the close interval $[0,1]$ of the real axis onto the parallelepiped (\ref{eq:02}) unambiguously and continuously.
@@
 
-It should be noted that if the convolution $F(\lambda,y)$ satisfies the Lipschitz condition then the reduced function $\varphi(y(x)) = F(\lambda,y(x))$ meets the H{\" o}lder condition
+It should be noted that if the convolution $F(\lambda,y)$ satisfies the Lipschitz condition then the reduced function $\varphi(y(x)) = F(\lambda,y(x))$ meets the H{\" o}lder inequality
 \begin{equation}
@@
 
+\textbf{Optimization.} This stage consists in solving a family of problems (\ref{eq:07}) corresponding to a finite set of weight coefficients $\lambda$.
+After minimization of all the problems from this family we get an approximation of the Pareto set.
+To solve problems (\ref{eq:07}), the information-statistical global search algorithm \cite{ML_MCO_2023,Gergel2019_2,Gergel2018,GergelKozinov2020,Strongin2000,Sergeyev2013} was used which provides efficient optimization of H{\" o}lderian functions.
+Computational scheme of the algorithm (call it briefly GSA) consists in the following.
+
+Each iteration of the search consists of one trial execution. The term ``trial'' means evaluation of the minimized function $\varphi(x)$ from (\ref{eq:07}) at some point.
+
+The first two trials are to be carried out at the boundary points of the interval $[0,1]$, i.e., at points $x^0=0$, $x^1=1$.
+
+Let $k$, $k \geq 2$, search iterations have been executed at trial points $x^0, x^1,\dots,x^{k-1}$ with objective function values, where $z^0, z^1,\dots,z^{k-1}$, $z^i=\varphi(x^i)$, $1 \leq i \leq k-1$.
+Then the choice of the new trial point $x^k$ is determined according to the following rules.
+
+\textit{Rule 1.} Renumber by subscripts all trial points in increasing order
 \begin{equation}
@@
 \end{equation}
+and juxtapose to them the values $z_i=\varphi(x_i)$, $1 \leq i \leq k-1$.
+
+\textit{Rule 2.} Calculate the current estimation of the H{\" o}lder constant
 
@@
 			\end{cases} ,
-		M = \max_{1 \leq i \leq k} \frac{| z_i - z_{i-1}|}{\varrho_i}, \\
+		M = \max_{1 \leq i \leq k-1} \frac{| z_i - z_{i-1}|}{\varrho_i}, \\
 		z_i = \varphi( x_i ), \varrho_i=\sqrt[N]{x_i-x_{i-1}}
@@
 \end{equation}
+where $r$, $r>1$ is the \textit{reliability} parameter of the algorithm.
+
+\textit{Rule 3.} For each subinterval $(x_{i-1} ,x_i)$, $1 \leq i \leq k-1$, compute the value
 
@@
     \label{eq:11}
-    R(j) = \varrho_j + \frac{(z_i-z_{i-1})^2}{m^2 \varrho_i} - \frac{2 (z_i+z_{i-1})}{m}, 1 \leq i \leq k.
+    R(i) = \varrho_i + \frac{(z_i-z_{i-1})^2}{m^2 \varrho_i} - \frac{2 (z_i+z_{i-1})}{m}, 1 \leq i \leq k-1.
+\end{equation}
+called \textit{characteristic} of the subinterval $(x_{i-1} ,x_i)$, $1 \leq i \leq k-1$.
+
+\textit{Rule 4.} Choose the subinterval $(x_{t-1} ,x_t)$ with maximal characteristic $R(t)$
+\begin{equation}
+    \label{eq:12}
+    R(t) = max_{1 \leq i \leq k-1} {R(i)}.
 \end{equation}
 
-\begin{equation}
-    \label{eq:12}
-    R(t) = argmax_{1 \leq i \leq k} {R(i)}.
-\end{equation}
-
+\textit{Rule 5.} Accept
 \begin{equation}
     \label{eq:13}
-    x^{k+1} = \frac{x_t + x_{t-1}}{2} - sign(z_t - z_{t-1}) \frac{1}{2r} \left(\frac{|z_l - z_{l-1}|}{m} \right)^N,
+    x^{k} = \frac{x_t + x_{t-1}}{2} - sign(z_t - z_{t-1}) \frac{1}{2r} \left(\frac{|z_l - z_{l-1}|}{m} \right)^N,
 \end{equation}
+as the point for the next trial.
 
+The process of finding a solution to the current problem (\ref{eq:07}) is terminated when the condition
 \begin{equation}
@@
 \end{equation}
+is met where $\varepsilon > 0$ is a preassigned search accuracy.
 
@@
 \label{sec:4}
+
+To enhance the efficiency of the Pareto set finding, several GSA modifications are used in the described approach.
+
+During minimization of one problem (\ref{eq:07}) with corresponding weight vector $\lambda$, the trial points $x^0, x^1,\dots,x^{k-1}$ from the interval $[0,1]$ generate points $y^0, y^1,\dots,y^{k-1}$ in the domain $D$, where $y^i=y(x^i)$, $1 \leq i \leq k-1$, and $y(x)$ is the Peano mapping.
+After computing values $f^i_j=f_j(y^i)$, $1 \leq i \leq k-1$, $1 \leq j \leq s$, of the partial criteria at points $y^i$ one can organize the set
+\begin{equation}
+    \label{eq:15}
+    \Omega_K=\{(x^q,y^q,f^q_1,\dots,f^q_s)^T: 0 \leq q \leq K\}.
+\end{equation}
+called \textit{Search State Matrix} (SSM).
+
+Solving new problems (\ref{eq:04}) allows one to replenish SSM with new elements and to save the additional information on the investigated problem (\ref{eq:02a}).
+This information can accelerate the minimization process for problems (\ref{eq:07}) and, as a consequence, for the initial multicriteria problem.
+
+One of the ways how this can be done is connected with the modification of the starting iterations of GSA.
+Namely, if $P \geq 1$ problems (\ref{eq:07}) have been already solved and the matrix (\ref{eq:15}) has been formed, then as initial trial coordinates, the points $x^q$, $1 \leq q \leq K$, can be used with values $z^q = F(\lambda,y^q)$.
+Moreover, the computation of the convolution (\ref{eq:05}) does not require the new calculation of criteria values at these points, because these values are taken from the matrix (\ref{eq:15}).
+After that, one can go to the basic iteration Rules 1-5.
+Such the scheme reduces significantly the number of criteria computations when jointly solving the family of problems (\ref{eq:04}) with convolution (\ref{eq:05}) for different weight coefficients $\lambda$ \cite{ML_MCO_2023,Gergel2018,GergelKozinov2020}.
+
+The second approach also aims to further decreasing the number of trials required to find the Pareto set.
+It suggests to change the form of the characteristics (\ref{eq:11}) in the GSA scheme.
+It is proposed to use the characteristic
+\begin{equation}
+    \label{eq:16}
+    R(i) = R_{gsa} (i) +  \alpha R_{ps} (i).
+\end{equation}
+instead of the characteristic (\ref{eq:11}) where $R_{gsa}(i)$ coincides with the expression (\ref{eq:11}) and the additional term $R_{ps}(i)$ depends on closeness of the points $y(x_{i-1})$, $y(x_{i})$ to the Pareto set, namely, the closer these points to the Pareto set, the greater the value $R_{ps}(i)$.
+
+The parameter $\alpha > 0$ allows changing the degree of influence of the $R_{ps}(i)$ on the choice of the next trial point.
+If $\alpha$ is small, the algorithm spends more trials to find the optimal solution to the current optimization problem.
+When $\alpha$ is large, the algorithm chooses the subintervals providing the closeness to the set of efficient points.
+In a certain sense, $R_{ps}(i)$ can be considered as a penalty for computations outside the Pareto set.
+
+For computation of $R_{ps}(i)$ we propose to apply the machine learning methods.
+It should be noted that in classical combinations of machine learning and optimization methods, the latter are used as an auxiliary internal techniques that improve the machine learning functioning.
+In our approach, on the contrary, machine learning methods are a part of the optimization algorithms enhancing their work.
+
+In the paper \cite{ML_MCO_2023} we proposed an algorithm for calculation of the penalty term $R_{ps}(i)$ on the base of a hyperplane separating the Pareto and other trial points (see Subsection \ref{ssec:41}).
+The changes in the characteristics led to a significant increase in the efficiency of the algorithm, but the question of the effectiveness of the proposed penalty term in problems where the Pareto set is non-convex and has a complex form, as well as the number of criteria is more than two, remained open.
+
+In the framework of this study, we propose several new ways to calculate the penalty term (see Subsection \ref{ssec:42}) and investigate their efficiency in solving a series of optimization problems (see Section \ref{sec:5}).
 
@@
 \label{ssec:42}
-
-\begin{equation}
-    \label{eq:15}
-    \Omega_k=\{\mu^i = (y^i,f^i=f(y^i))^T: 0 \leq i \leq k\}.
-\end{equation}
-
-\begin{equation}
-    \label{eq:16}
-    R(i) = R_{gsa} (i) +  \alpha R_{ps} (i).
-\end{equation}
 

### 2024-06-28 2024/OPTIMA/OPTIMA_2024_MCO.tex
check current part

@@
 
-Each iteration of the search consists of one trial execution. The term ``trial'' means evaluation of the minimized function $\varphi(x)$ from (\ref{eq:07}) at some point.
+Each iteration of the search consists of one trial execution. The term ``trial'' means evaluation of the objective function $\varphi(y(x))$ from (\ref{eq:07}) at some point.
 
@@
 
-Let $k$, $k \geq 2$, search iterations have been executed at trial points $x^0, x^1,\dots,x^{k-1}$ with objective function values, where $z^0, z^1,\dots,z^{k-1}$, $z^i=\varphi(x^i)$, $1 \leq i \leq k-1$.
+Let $k$, $k \geq 2$, search trials have been executed at the points $x^0, x^1,\dots,x^{k-1}$ with the objective function values $z^i=\varphi(y(x^i))$, $1 \leq i \leq k-1$.
 Then the choice of the new trial point $x^k$ is determined according to the following rules.
@@
     \label{eq:09}
-    x_0 = 0 < x_1 < \dots < x_i < \dots < x_{k} = 1.
+    0 = x_0 < x_1 < \dots < x_i < \dots < x_{k-1} = 1.
 \end{equation}
-and juxtapose to them the values $z_i=\varphi(x_i)$, $1 \leq i \leq k-1$.
+and juxtapose to them the values $z_i=\varphi(y(x_i))$, $1 \leq i \leq k-1$.
 
@@
 					 1, & M = 0 
-				\end{matrix} 
-			\end{cases} ,
+				\end{matrix} \; , 
+			\end{cases}
 		M = \max_{1 \leq i \leq k-1} \frac{| z_i - z_{i-1}|}{\varrho_i}, \\
-		z_i = \varphi( x_i ), \varrho_i=\sqrt[N]{x_i-x_{i-1}}
+		z_i = \varphi( y(x_i) ), \varrho_i=\sqrt[N]{x_i-x_{i-1}}
 		\end{matrix}
 \end{equation}
-where $r$, $r>1$ is the \textit{reliability} parameter of the algorithm.
+where $r>1$ is the \textit{reliability} parameter of the algorithm.
 

### 2024-06-28 2024/OPTIMA/OPTIMA_2024_MCO.tex
OPTIMA part 1 fixed

@@
 %
-\title{Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriteria problems\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.
+\title{Comparative efficiency of machine learning models for enhancing algorithms in solving multiextremal multicriteria problems\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.
 FSWR-2023-0034, and by by the Research and Education Mathematics Centre ``Mathematics for Future Technologies'', project no.
@@
 \begin{abstract}
-To solve a multicriteria optimization problem, it is necessary to find a whole set of independent variables corresponding to non-dominated criteria values (Pareto set).
+Solving a multicriteria optimization problem requires finding a whole set of independent variables corresponding to non-dominated criteria values (Pareto set).
 The complexity of these problems increases significantly in the case when the criteria are multiextremal.
-The paper presents several variations of the global optimization algorithm intended for solving black-box multiobjective problems with multiextremal criteria.
-The developed methods are based on a combination of the information-statistical approach and machine learning procedures aimed at increasing the efficiency of constructing the Pareto set.
+The paper presents several variations of the global optimization algorithm intended for solving ``black-box'' multiobjective problems with multiextremal criteria.
+The proposed methods are based on a combination of the information-statistical approach and machine learning procedures aimed at increasing the efficiency of constructing the Pareto set.
 As a novelty, the paper describes a new approach to using machine learning models within the algorithm of multicriteria optimization.
-The effectiveness of different versions of the developed algorithm is estimated on the base of  a representative computational experiment.
+The effectiveness of different versions of the proposed algorithm is estimated on the base of a representative computational experiment.
 
@@
 
-Problems of optimal choice are widespread in many areas of activity.
-As a rule, in complicated decision making models there are several functions to be optimized which reflect the decision making objectives and those are, in general case, contradictory.
-In these circumstances, the notion of optimal solution is more complicated than in single criterion (scalar) statement formulated as mathematical programming problem because there exist incomparable decisions.
-In this situation under introduced definition of domination, the set of non-dominated variants (Pareto set) is taken as the solution of the multiobjective problem.
+Optimal choice problems are very common in many areas of activity.
+As a rule, in complex decision-making models there are several functions to be optimized which reflect the decision-making objectives that are in the general case in conflict with each other.
+In these circumstances, the notion of optimal solution is more complicated than in a single criterion (scalar) statement formulated as a mathematical programming problem because there exist incomparable decisions.
+In this situation, with the introduced definition of domination, the set of non-dominated variants (Pareto set) is taken as the solution of the multiobjective problem.
 
-In our study we deal with methods for solving the most complex class of multicriteria problems where criteria are considered as ``black-box'' functions and are multiextremal \cite{Miettinen1999,Ehrgott2005,Pardalos2017,Strongin2000,Sergeyev2013}.
+In our study, we deal with methods for solving the most complex class of multicriteria problems where criteria are considered as ``black-box'' functions and are multiextremal \cite{Miettinen1999,Ehrgott2005,Pardalos2017,Strongin2000,Sergeyev2013}.
 
-It should be noted that problems of multiextremal optimization are among the most complicated problems of mathematical programming.
-Unlike the local optimization where an optimal point has to be juxtaposed with other points only in its region of attraction, the global solution must be better w.r.t.
-all points in the feasible domain.
-If to solve such problems by grid search methods, the amount of computations will rise exponentially when increasing the number of parameters (variables of the problem).
+It should be noted that multiextremal optimization problems are among the most complicated problems of mathematical programming.
+Unlike the local optimization where an optimal point has to be juxtaposed with other points only in its region of attraction, the global solution must be better with respect to all points in the feasible domain.
+If one tries to solve such problems by grid search methods, the amount of computations will rise exponentially when increasing the number of parameters (variables of the problem).
 
-For solving the multiextremal problems the more effective algorithms are used which can be divided into two key classes, namely, into deterministic (see, e.g., \cite{Evtushenko2014,Gergel2018,GergelKozinov2020,Paulavicius2020,Jones2021}) and meta-heuristic (see, e.g., \cite{Battiti2009,Gendreau2010,Eiben2015}) methods .
+For solving multiextremal problems, more effective algorithms are used which can be divided into two key classes, namely, into deterministic(see, e.g., \cite{Evtushenko2014,Gergel2018,GergelKozinov2020,Paulavicius2020,Jones2021}) and meta-heuristic (nature-inspired) methods (see, e.g., \cite{Battiti2009,Gendreau2010,Eiben2015}).
 
-Meta-heuristic algorithms model the behavior of living organisms or nature processes \cite{Deb2002,Durillo2010,Mostaghim2007,NDG09,RC05,ZLT01}.
+Meta-heuristic algorithms model the behavior of living organisms or natural processes \cite{Deb2002,Durillo2010,Mostaghim2007,NDG09,RC05,ZLT01}.
 As a rule, these methods are simple enough in implementation and can demonstrate good efficiency for unimodal objective functions and multiextremal problems with a small number of local optima; however, they do not ensure finding the best solution.
 
-Deterministic algorithms guarantee convergence to global optimum, but their computational schemes are more complex.
+Deterministic algorithms guarantee convergence to the global optimum, but their computational schemes are more complex.
 However, advanced deterministic algorithms outperform metaheuristic ones in terms of quality \cite{Kvasov2018,Sergeyev2018}.
 
-One of the qualitative deterministic methods for solving multiextremal problems is the  information-statistical algorithm of global search \cite{Strongin2000,Sergeyev2013}.
+One of the proven deterministic methods for solving multiextremal problems is the information-statistical algorithm of global search \cite{Strongin2000,Sergeyev2013}.
 This algorithm proposed initially for solving scalar problems was successfully developed for analyzing multiobjective statements \cite{ML_MCO_2023,Gergel2018,GergelKozinov2020}.
 
-In the paper \cite{ML_MCO_2023} we proposed an approach to solving multicriteria optimization problems combining the classic information-statistical global search algorithm with machine learning methods.
-In the framework of the approach a problem of multicriteria optimization was transformed to a series of scalar optimization subproblems.
-In the course of solving scalar subproblems, on the base of machine learning methods a hyperplane was built and gradually refined which separated points of effective and ineffective solutions in the criteria space.
-Additional information connected with the separating hyperplane was embedded in the computational scheme of the optimization algorithm and, as a result, the number of trials (criteria evaluations) in domains with ineffective solutions was significantly decreased.
-The experimental results showed that in comparison with a number of heuristic algorithms, the search for effective solutions was several times faster.
+In our publication \cite{ML_MCO_2023}, we proposed an approach to solving multicriteria optimization problems combining the classic information-statistical global search algorithm with machine learning methods.
+In the framework of this approach, the problem of multicriteria optimization was transformed into a series of scalar optimization subproblems.
+In the course of solving scalar subproblems, a hyperplane separating the points of efficient and inefficient solutions in the criteria space was built and gradually refined using machine learning methods.
+Additional information related to the separating hyperplane was embedded in the computational scheme of the optimization algorithm and, as a result, the number of trials (criteria evaluations) in domains with inefficient solutions was significantly decreased.
+The experimental results indicate that, compared to a number of heuristic algorithms, the search for efficient solutions is several times faster.
+
 At the same time, this approach can be improved because utilizing only the separating plane limits the class of machine learning techniques which can be applied for enhancing the global search algorithm.
 
-In the framework of our study we propose a new approach to separating search trial points in the criteria space based on the probability that reflects the belonging of the trial point to the Pareto set.
-The most machine learning models allow obtaining this metric which makes it possible to choose an algorithm for building a machine learning model depending on the features of the problem being solved.
+In the framework of our study we propose a new approach to separating search trial points in the criteria space based on the probability that shows whether a trial point belongs to the Pareto set.
+Most machine learning models allow obtaining this metric which makes it possible to choose an algorithm for building a machine learning model depending on the features of the problem being solved.
 
@@
 Section \ref{sec:2} describes the mathematical statement of the problem to be studied.
-Section \ref{sec:3} is devoted to a brief description of the basic algorithms used for solving problems of multicriteria optimization.
+Section \ref{sec:3} presents a brief description of the basic algorithms used for solving problems of multicriteria optimization.
 Section \ref{sec:4} contains information on the techniques applied for reducing the number of trials during the optimization process.
@@
 
-Let $k$, $k \geq 2$, search trials have been executed at the points $x^0, x^1,\dots,x^{k-1}$ with the objective function values $z^i=\varphi(y(x^i))$, $1 \leq i \leq k-1$.
+Let $k$, $k \geq 2$, search trials have been executed at the points $x^0, x^1,\dots,x^{k-1}$ with the objective function values $z^i=\varphi(y(x^i))$, $0 \leq i \leq k-1$.
 Then the choice of the new trial point $x^k$ is determined according to the following rules.
@@
 \end{equation}
-and juxtapose to them the values $z_i=\varphi(y(x_i))$, $1 \leq i \leq k-1$.
+and juxtapose to them the values $z_i=\varphi(y(x_i))$, $0 \leq i \leq k-1$.
 
@@
     \label{eq:13}
-    x^{k} = \frac{x_t + x_{t-1}}{2} - sign(z_t - z_{t-1}) \frac{1}{2r} \left(\frac{|z_l - z_{l-1}|}{m} \right)^N,
+    x^{k} = \frac{x_t + x_{t-1}}{2} - sign(z_t - z_{t-1}) \frac{1}{2r} \left(\frac{|z_t - z_{t-1}|}{m} \right)^N,
 \end{equation}
@@
 
-During minimization of one problem (\ref{eq:07}) with corresponding weight vector $\lambda$, the trial points $x^0, x^1,\dots,x^{k-1}$ from the interval $[0,1]$ generate points $y^0, y^1,\dots,y^{k-1}$ in the domain $D$, where $y^i=y(x^i)$, $1 \leq i \leq k-1$, and $y(x)$ is the Peano mapping.
-After computing values $f^i_j=f_j(y^i)$, $1 \leq i \leq k-1$, $1 \leq j \leq s$, of the partial criteria at points $y^i$ one can organize the set
+During minimization of one problem (\ref{eq:07}) with corresponding weight vector $\lambda$, the trial points $x^0, x^1,\dots,x^{k-1}$ from the interval $[0,1]$ generate points $y^0, y^1,\dots,y^{k-1}$ in the domain $D$, where $y^i=y(x^i)$, $0 \leq i \leq k-1$, and $y(x)$ is the Peano mapping.
+After computing values $f^i_j=f_j(y^i)$, $0 \leq i \leq k-1$, $1 \leq j \leq s$, of the partial criteria at points $y^i$ one can organize the set
 \begin{equation}
@@
 One of the ways how this can be done is connected with the modification of the starting iterations of GSA.
-Namely, if $P \geq 1$ problems (\ref{eq:07}) have been already solved and the matrix (\ref{eq:15}) has been formed, then as initial trial coordinates, the points $x^q$, $1 \leq q \leq K$, can be used with values $z^q = F(\lambda,y^q)$.
+Namely, if $P \geq 1$ problems (\ref{eq:07}) have been already solved and the matrix (\ref{eq:15}) has been formed, then as initial trial coordinates, the points $x^q$, $0 \leq q \leq K$, can be used with values $z^q = F(\lambda,y^q)$.
 Moreover, the computation of the convolution (\ref{eq:05}) does not require the new calculation of criteria values at these points, because these values are taken from the matrix (\ref{eq:15}).

### 2024-06-28 2024/OPTIMA/OPTIMA_2024_MCO.tex
fix \max
OPTIMA part 5

@@
     \label{eq:12}
-    R(t) = max_{1 \leq i \leq k-1} {R(i)}.
+    R(t) = \max_{1 \leq i \leq k-1} {R(i)}.
 \end{equation}
@@
 is met where $\varepsilon > 0$ is a preassigned search accuracy.
+
+%Theoretical justification of this algorithm is provided in \cite{Strongin2000}.
 
@@
 
-\subsection{Section 41}
+\subsection{Constructing the penalty term on the base of separating hyperplane}
 \label{ssec:41}
 
-\subsection{Section 42}
-\label{ssec:42}
+In the course of solving the MCO problem, with each change in the $\lambda$ coefficients, and therefore, the transition to a new optimization problem from (\ref{eq:04}), all obtained trial points in SSM (\ref{eq:15}) are divided into two classes.
+The first class includes points belonging to the Pareto set.
+All other trial points belong to the second class.
+The number of points that are not related to the Pareto domain is usually much larger, and therefore, the classes are not balanced.
+For the better functioning of machine learning algorithms, for each class the weights corresponding to the significance of the trial points are set if possible.
+Weights for classes are set at the beginning of the algorithm running.
+Subsequently, a separating  hyperplane is built between the two classes.
+To construct the hyperplane, the SVC algorithm with the linear kernel from the scikit-learn library \cite{SVM_2000,PROB_2004} is used.
 
+After the construction of the separating hyperplane, for each trial point the $d_i$ distance is calculated and normalized according to the following expression:
 \begin{equation}
@@
 \end{cases}, 
-1 \leq i \leq k,
+0 \leq i \leq k-1,
 \end{equation}
+where $d_i$ is the distance from the $i$-th point to the hyperplane, $d_{max}$, $d_{min}$ are maximal and minimal distances to the hyperplane.
+
+The value $R_{ps}(i)$ from (\ref{eq:16}) for the interval $(x_{i-1},x_{i})$ is calculated as the sum of normalized distances (\ref{eq:17}) to points $x_i$ and $x_{i-1}$.
+Let's denote this version of GSA as ML\_MGSA\_Dist.
+
+\subsection{Building the penalty term on the base of probabilities being derived from machine learning models}
+\label{ssec:42}
+
+For creating the penalty term $R_{ps}(i)$ from (\ref{eq:16}) that corresponds better to the estimation of the Pareto set, the following new ideas are suggested.
+
+\begin{enumerate}
+	\item Probabilistic approach with application of machine learning models.   
+	\item Automatic adjustment of class weights when building a machine learning model.
+	\item Transformation of machine learning probabilities.
+\end{enumerate}
+
+As part of the proposed approach, the machine learning model is built according to the rules similar to the construction of the SVC model from section \ref{ssec:41}.
+As before, all trial points are divided into two classes -- points of the Pareto set and points that do not belong to the Pareto set.
+The classifier is trained on the selected classes.
+Note that the classifier models of most machine learning libraries can return not only the class label, but also the probability of relevance to the class.
+Unlike the approach described in Subsection \ref{ssec:41} the characteristic $R_{ps}(i)$ is computed as the average of probabilities of situations when the points $x_i$ and $x_{i-1}$ from interval $(x_{i-1},x_{i})$ belong to the Pareto set.
+
+A significant advantage of this approach is the possibility to easily replace both the machine learning model itself and its learning rules.
+In the framework of the study, several variants of building models that potentially improve the factor $R_{ps}(i)$ to beter solve the MCO problems.
+
+The first direction is to change the method of building the classifier. Changing the classifier can be useful if the Pareto set is complex, for instance, non-convex.
+For comparison the SVM models with linear kernel, polynomial kernel, and RBF function-based kernel are considered.
 
 \begin{figure}
-\center
+\centering
 \includegraphics[width=0.8\textwidth]{fig1.png}
-\label{fig1}
-\caption{A figure caption.} 
+\caption{Heatmap of the function $R_ps(i)$ in the criteria space for algorithms with different machine learning models.} 
+\label{fig:1}
 \end{figure}
 
+Fig.
+\ref{fig:1} shows the behavior of the characteristic $R_{ps}(i)$ in the criteria space for different models.
+Red points are related to the Pareto set, orange those to the class  where points are not effective ones.
+
+The color background displays the behavior of the penalty $R_{ps}(i)$ as a function of the criteria $f1$ and $f2$.
+Fig.
+\ref{fig:1}(a) shows the function $R_{ps}(i)$ for the method ML\_MGSA\_Dist.
+Fig.
+\ref{fig:1}(b), \ref{fig:1}(c), and \ref{fig:1}(d) demonstrate $R_{ps}(i)$ for the different cases of the probabilistic models.
+Fig.
+\ref{fig:1}(b) corresponds to the model with linear kernel (ML\_MGSA\_Prob\_Linear), Fig.
+\ref{fig:1}(c) reflects results for the model with polynomial kernel (ML\_MGSA\_Prob\_Poly), and Fig.
+\ref{fig:1}(d) is related to RBF-based kernel (ML\_MGSA\_Prob\_RBF).
+
+The number of trials carried out in each experiment is indicated in parentheses after the name of the method.
+
+The figure shows that the penalty function with kernels other than linear describes the Pareto set well, but, may be, penalty is not sufficient to reduce the number of trials of the global search algorithm.
+Two modifications can be proposed to solve this problem.
+
+First, when changing the optimization problem from set (\ref{eq:07}), a new weight calculation scheme for each of two classes of trial points is proposed.
+In this scheme, the weights of each class are calculated inversely proportional to the ratio of the number of all points to the number of points belonging to the Pareto set.
+
+Second, the probabilities undergo the transformation
 \begin{equation}
     \label{eq:18}
-p'_i= \frac{ \log (p_i) - \log (p_{min})}{ \log (p_{max}) - \log (p_{min})} , 1 \leq i \leq k,
+p'_i= \frac{ \log (p_i) - \log (p_{min})}{ \log (p_{max}) - \log (p_{min})} , 0 \leq i \leq k-1,
 \end{equation}
+where $p_i$ is a probability of the event that the point $x_i$ belongs to the Pareto set, $p_{min}$, $p_{max}$ are minimum and maximum values of such probabilities.
+After transformation the function $R_{ps}(i)$ is calculated, as earlier, as the average value of probabilities of events in which the points $x_i$ and $x_{i-1}$ from interval $(x_{i-1}, x_i)$ belong to the Pareto set.
+
+Computational experiments were conducted for each method and each model in order to assess the effectiveness of using these techniques in the global search algorithm.
+Fig.
+\ref{fig:2} demonstrates the visualization of $R_{ps}(i)$ similar to Fig.
+\ref{fig:1}.
 
 \begin{figure}
-\center
+\centering
 \includegraphics[width=0.8\textwidth]{fig2.png}
-\label{fig2}
-\caption{A figure caption.} 
+\caption{Heatmap of the function $R_{ps}(i)$ in the criteria space for algorithms with different techniques of building models after transformation (\ref{eq:18}).} 
+\label{fig:2}
 \end{figure}
+
+It is easy to notice that new penalty functions (see Fig.
+\ref{fig:2}.
+(b)-(d)) describe better the Pareto set than functions built on the base of separating hyperplane.
+A detailed comparison of approaches to different computation of the characteristic (\ref{eq:16}) is given in Section \ref{sec:5}.
 
@@
 \includegraphics[width=\textwidth]{fig3.png}
-\label{fig3}
+\label{fig:3}
 \caption{A figure caption.} 
@@
 \includegraphics[width=\textwidth]{fig4.png}
-\label{fig4}
+\label{fig:4}
 \caption{A figure caption.} 
@@
 \includegraphics[width=\textwidth]{fig5.png}
-\label{fig5}
+\label{fig:5}
 \caption{A figure caption.} 

### 2024-06-28 2024/OPTIMA/OPTIMA_2024_MCO.tex
minor changes

@@
 \end{equation}
-where $d_i$ is the distance from the $i$-th point to the hyperplane, $d_{max}$, $d_{min}$ are maximal and minimal distances to the hyperplane.
+where $d_i$ is the distance from the $i$-th point to the hyperplane, $d_{max}$, $d_{min}$ are maximum and minimum distances to the hyperplane.
 
@@
 
-For creating the penalty term $R_{ps}(i)$ from (\ref{eq:16}) that corresponds better to the estimation of the Pareto set, the following new ideas are suggested.
-
+For creating the penalty term $R_{ps}(i)$ from (\ref{eq:16}) that corresponds better to the estimation of the Pareto set, the following new ideas are suggested: 
 \begin{enumerate}
-	\item Probabilistic approach with application of machine learning models.   
-	\item Automatic adjustment of class weights when building a machine learning model.
+	\item Probabilistic approach with application of machine learning models;   
+	\item Automatic adjustment of class weights when building a machine learning model;
 	\item Transformation of machine learning probabilities.
@@
 
-The color background displays the behavior of the penalty $R_{ps}(i)$ as a function of the criteria $f1$ and $f2$.
+The color background displays the behavior of the penalty $R_{ps}(i)$ as a function of the criteria $f_1(y)$ and $f_2(y)$.
 Fig.
@@
 Fig.
-\ref{fig:1}(b), \ref{fig:1}(c), and \ref{fig:1}(d) demonstrate $R_{ps}(i)$ for the different cases of the probabilistic models.
+\ref{fig:1}(b,c,d) demonstrate $R_{ps}(i)$ for the different cases of the probabilistic models.
 Fig.
@@
 \ref{fig:1}(d) is related to RBF-based kernel (ML\_MGSA\_Prob\_RBF).
-
 The number of trials carried out in each experiment is indicated in parentheses after the name of the method.
@@
 \ref{fig:1}.
-
 \begin{figure}
@@
 
-It is easy to notice that new penalty functions (see Fig.
-\ref{fig:2}.
-(b)-(d)) describe better the Pareto set than functions built on the base of separating hyperplane.
+Note that new penalty functions (see Fig.
+\ref{fig:2}(b,c,d)) describe better the Pareto set than functions built on the base of separating hyperplane.
 A detailed comparison of approaches to different computation of the characteristic (\ref{eq:16}) is given in Section \ref{sec:5}.

### 2024-06-29 2024/OPTIMA/OPTIMA_2024_MCO.tex
OPTIMA last part

@@
 %
+\usepackage{multirow}
 \usepackage{graphicx}
@@
 \label{sec:1}
-
-\cite{Miettinen1999,Ehrgott2005,Pardalos2017,ML_MCO_2023,Evtushenko2014,Deb2002,Durillo2010,Mostaghim2007,NDG09,RC05,ZLT01,Gergel2019_2,Gergel2018,GergelKozinov2020,Marler2004,Strongin2000,Sergeyev2013,SVM_2000,PROB_2004,iOpt_url,Grishagin2015_2}
 
@@
 
+Computations were conducted on the supercomputer ``Lobachevsky'' of the Nizhni Novgorod State University (operating system CentOS 7, control system SLURM).
+To obtain the executable program code the compilers Intel C++ 17.0.0 and python 3.9 were used.
+Computations were performed with the support of the package scikit-learn 0.24.2 and the optimization software iOpt \cite{iOpt_url}.
+
+In the first experiment 100 bi-criterial MCO problems have been solved. As the partial criteria, the functions \cite{ML_MCO_2023} 
 \begin{equation}
@@
 \end{equation}
+were taken  where $0 \leq y_1, y_2 \leq 1$, parameters $-1 \leq A_{ij},B_{ij},C_{ij},D_{ij} \leq 1$ are independent uniformly distributed random values.
+
+Computational experiments for techniques of Subsection \ref{ssec:42} which utilize probabilities of belonging to the Pareto set can be divided into 3 groups.
+\begin{enumerate}
+	\item 	Experiments with the algorithm ML\_MGSA\_Prob (fixed weights).
+	\item 	Solving MCO problems by the method ML\_MGSA\_Prob (Adjustive weights));
+	\item 	Experiments with changing weights in the course of ML\_MGSA\_Prob running and transformed probabilities (\ref{eq:18})  (ML\_MGSA\_Prob (Norm Log)).
+\end{enumerate}
+
+In each series of experiments for these optimization algorithms, the coefficients $\alpha$ from (\ref{eq:16}) changed in the range $[0.1, 0.9]$.
+Comparison with the algorithm MGSA that does not use machine learning methods (in the characteristic ($\alpha$) $\alpha = 0$) and with the algorithm ML\_MGSA\_Dist taking into account the distance to the separating hyperplane was carried out.
+The comparison was made on the base of  the calculated HV index characterizing quality of the obtained Pareto set approximation \cite{ML_MCO_2023,Evtushenko2014}.
+Further subsections include the results of only those calculations in which the $\alpha$ parameter provided the best quality of the solution found.
+Tables contain results of solving all the problems (\ref{eq:07}) with fixed accuracy $\varepsilon=0.01$ in the termination condition (\ref{eq:14}).
+Figures present results for a series of decreasing $\varepsilon$ which generates increasing numbers of trials.
+
+\textbf{Experiments with fixed weights of two trial classes.} The first series of experiments is aimed at comparison the algorithm ML\_MGSA\_Prob (Fixed weights) with results of running for methods MGSA and ML\_MGSA\_Dist.
+
+The following parameters are indicated in Table \ref{tab:1}: the kernel of the SVM model (column 2), parameter $\alpha$ (column 3), average values of the HV-index (column 4), the average number of trials spent by the algorithm (column 5), and the coefficient of decreasing the number of trials compared to ML\_MGSA (column 6).
+The factors $r$ and $\varepsilon$ from (\ref{eq:10}) and (\ref{eq:14}) were the same for all the algorithms.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+% \usepackage{graphicx}
+\begin{table}[ht]
+\centering
+\caption{Results of computational experiments}
+\label{tab:1}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{cccccc}
+\hline
+\textbf{}                                                                                  & \textbf{Model Kernel} & \textbf{$\alpha$} & \textbf{Avg.
+HV-index} & \textbf{Avg.
+iter.} & \textbf{Speedup iter.} \\ \hline
+ML\_MGSA                                                                                   & -                     & -                              & 114.91                 & 2269.82             & 1.00                   \\ \hline
+\multicolumn{6}{c}{\textbf{Experiments for the case of fixed weights}}                                                                                                                                                      \\ \hline
+ML\_MGSA\_Dist                                                                             & Linear                & 0.01                           & 114.67                 & 691.14              & \textbf{3.28}          \\
+\multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}ML\_MGSA\_Prob \\ (Fixed weights)\end{tabular}} & Linear                & 0.04                           & 114.72                 & 963.05              & 2.36                   \\
+                                                                                           & Poly                  & 0.07                           & 114.73                 & 871.82              & 2.60                   \\
+                                                                                           & RBF                   & 0.03                           & 114.80                 & 883.04              & 2.57                   \\ \hline
+\multicolumn{6}{c}{\textbf{Experiments for the case of adjusted weights}}                                                                                                                                                   \\ \hline
+ML\_MGSA\_Dist                                                                             & Linear                & 0.01                           & 114.67                 & 691.14              & \textbf{3.28}          \\
+\multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}ML\_MGSA\_Prob \\ (Adj.
+weights)\end{tabular}}  & Linear                & 0.04                           & 114.76                 & 1018.03             & 2.23                   \\
+                                                                                           & Poly                  & 0.08                           & 114.73                 & 844.66              & 2.69                   \\
+                                                                                           & RBF                   & 0.03                           & 114.76                 & 796.74              & 2.85                   \\ \hline
+\multicolumn{6}{c}{\textbf{Experiments for the case of adjusted weights and logarithmic probabilities}}                                                                                                                              \\ \hline
+ML\_MGSA\_Dist                                                                             & Linear                & 0.01                           & 114.67                 & 691.14              & \textbf{3.28}          \\
+\multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}ML\_MGSA\_Prob \\ (Log Norm)\end{tabular}}      & Linear                & 0.05                           & 114.72                 & 711.44              & 3.19                   \\
+                                                                                           & Poly                  & 0.04                           & 114.79                 & 854.82              & 2.66                   \\
+                                                                                           & RBF                   & 0.04                           & 114.74                 & 704.99              & 3.22                   \\ \hline
+\multicolumn{6}{c}{\textbf{Experiments with 4-criterial MCO problems}}                                                                                                                                                      \\ \hline
+ML\_MGSA\_Dist                                                                             & Linear                & 0.01                           & 7751.43                & 1925.6              & 1.00                   \\
+\multirow{3}{*}{ML\_MGSA\_Prob (Log Norm)}                                                 & Linear                & 0.08                           & 7729.71                & 1760.5              & 1.09                   \\
+                                                                                           & Poly                  & 0.08                           & 7796.2                 & 4121.7              & 0.47                   \\
+                                                                                           & RBF                   & 0.08                           & 7786.48                & 1364.6              & \textbf{1.41}          \\ \hline
+\end{tabular}%
+}
+\end{table}
+
+Fig. \ref{fig:3} shows the change of the average HV-index as a function of the number of trials.
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{fig3.png}
+\caption{Comparison of HV-indices for the case of fixed weights.} 
+\label{fig:3}
+\end{figure}
+
+Compared to MGSA, the other algorithms reduce significantly the number of iterations in conditions when HV-indices are similar.
+However, the iteration number when using the models with probabilities (see last 3 rows of Table \ref{tab:1} at section ``Experiments for the case of fixed weights'') is greater than for the algorithm based on the separating hyperplane.
+
+\textbf{Experiments with tuned weights in the course of optimization.} In this group of experiments the algorithm ML\_MGSA\_Prob (adj.
+weights) changed weights of trial classes before starting a new subproblem (\ref{eq:07}).
+Table \ref{tab:1} and Fig.
+\ref{fig:4} for these experiments are organized similarly to the first series of experiments.
 
 \begin{figure}
-\includegraphics[width=\textwidth]{fig3.png}
-\label{fig:3}
-\caption{A figure caption.} 
+\centering
+\includegraphics[width=0.8\textwidth]{fig4.png}
+\caption{HV-indices comparison  in the case of adjusted weights} 
+\label{fig:4}
 \end{figure}
 
+The presented data on methods using probabilities, as in the previous series of experiments, show significantly better results in the number of iterations compared to an algorithm without machine learning techniques.
+Nevertheless, from the results obtained, one cannot unequivocally say about a significant improvement or deterioration of these results compared to experiments in which the weights were set fixed.
+
+\textbf{Experiments with adjusted weights and logarithmic probabilities.} In the third group of experiments, in addition to changing the weights during the algorithm’s running, the transformation (\ref{eq:18}) was used in techniques with probability calculation (methods ML\_MGSA\_Prob (Log Norm)).
+Table \ref{tab:1} and Fig.
+\ref{fig:5} are similar to those in the previous experiments.
+
 \begin{figure}
-\includegraphics[width=\textwidth]{fig4.png}
-\label{fig:4}
-\caption{A figure caption.} 
+\centering
+\includegraphics[width=0.8\textwidth]{fig5.png}
+\caption{Comparison of  HV-indices for methods with probabilities (\ref{eq:18})} 
+\label{fig:5}
 \end{figure}
 
+From the results of the experiment, the following conclusions can be drawn.
+Among the methods applying probabilistic techniques, the method with the transformation (\ref{eq:18}) demonstrates the best results.
+In the case of the SVM model with the linear kernel and with the RBF kernel, the results under increasing accuracy are comparable to the results obtained for ML\_MGSA\_Dist.
+At the same time, the techiques with probabilities enable changing easier the machine learning models.
+
+\textbf{Experiments with 4-criterial MCO problems.} In the last series of experiments, 10 MCO problems with 4 criteria were solved that form a Pareto set of complex structure.
+As in previous subsection, the functions of the family (\ref{eq:19}) were used as partial criteria.
+In the course of previous experiments, it was found that the closest result to ML\_MGSA\_Dist is given by the algorithm ML\_MGSA\_Prob (Log Norm) based on the technique (\ref{eq:18}).
+Taking into account this fact, we limited ourselves to comparing the results of these two algorithms.
+The results are presented in Table \ref{tab:1} and Fig.
+\ref{fig:6}
+
 \begin{figure}
-\includegraphics[width=\textwidth]{fig5.png}
-\label{fig:5}
-\caption{A figure caption.} 
+\centering
+\includegraphics[width=0.8\textwidth]{Fig6.png}
+\caption{Efficiency comparison in the 4-critarial case} 
+\label{fig:6}
 \end{figure}
+
+The results of the experiments allow us to conclude that the method ML\_MGSA\_Prob (Log Norm) in this series of experiments can significantly reduce the number of trials with a similar value of HV-index compared to ML\_MGSA\_Dist.
+The best results were achieved using models with Linear and RBF kernels, while the latter model demonstrates better quality than the version of the algorithm with the technique \cite{ML_MCO_2023} based on distances to the hyperplane.
+
+\section{Conclusion}
+
+The paper proposes a new modification of the global search algorithm focused on solving problems of multicriteria optimization.
+The algorithm in question is based on criteria scalarization and solving a series of multiextremal scalar problems.
+
+Acceleration of the algorithm (in terms of reducing the number of search trials required to evaluate the Pareto set) is achieved through the application of machine learning methods.
+Namely, when using the SVM method, trial points in the criteria space are classified (each point is/is not an effective solution), on the basis of which a numerical estimate of the probability of belonging to the Pareto set is built.
+Polynomial and radial basis function kernels in SVM allows one to adequately approximate Pareto front of a complex structure.
+
+Solving several series of MCO problems with complex multiextremal criteria has demonstrated that the proposed approach to evaluating the set of effective solutions is promising.
+Further research aims to develop the proposed approach by consideration of other machine learning models (in particular, artificial neural networks) to better classify trials and approximate the Pareto set.
 

### 2024-06-29 2024/OPTIMA/OPTIMA_2024_MCO.tex
OPTIMA fixed bugs

@@
 In each series of experiments for these optimization algorithms, the coefficients $\alpha$ from (\ref{eq:16}) changed in the range $[0.1, 0.9]$.
-Comparison with the algorithm MGSA that does not use machine learning methods (in the characteristic ($\alpha$) $\alpha = 0$) and with the algorithm ML\_MGSA\_Dist taking into account the distance to the separating hyperplane was carried out.
+Comparison with the algorithm MGSA that does not use machine learning methods (in the characteristic (\ref{eq:16}) $\alpha = 0$) and with the algorithm ML\_MGSA\_Dist taking into account the distance to the separating hyperplane was carried out.
 The comparison was made on the base of  the calculated HV index characterizing quality of the obtained Pareto set approximation \cite{ML_MCO_2023,Evtushenko2014}.
@@
 ML\_MGSA\_Dist                                                                             & Linear                & 0.01                           & 114.67                 & 691.14              & \textbf{3.28}          \\
-\multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}ML\_MGSA\_Prob \\ (Fixed weights)\end{tabular}} & Linear                & 0.04                           & 114.72                 & 963.05              & 2.36                   \\
+\multirow{3}{*}{ML\_MGSA\_Prob (Fixed weights)} & Linear                & 0.04                           & 114.72                 & 963.05              & 2.36                   \\
                                                                                            & Poly                  & 0.07                           & 114.73                 & 871.82              & 2.60                   \\
@@
 ML\_MGSA\_Dist                                                                             & Linear                & 0.01                           & 114.67                 & 691.14              & \textbf{3.28}          \\
-\multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}ML\_MGSA\_Prob \\ (Adj.
-weights)\end{tabular}}  & Linear                & 0.04                           & 114.76                 & 1018.03             & 2.23                   \\
+\multirow{3}{*}{ML\_MGSA\_Prob (Adj. weights)}  & Linear                & 0.04                           & 114.76                 & 1018.03             & 2.23                   \\
                                                                                            & Poly                  & 0.08                           & 114.73                 & 844.66              & 2.69                   \\
@@
 ML\_MGSA\_Dist                                                                             & Linear                & 0.01                           & 114.67                 & 691.14              & \textbf{3.28}          \\
-\multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}ML\_MGSA\_Prob \\ (Log Norm)\end{tabular}}      & Linear                & 0.05                           & 114.72                 & 711.44              & 3.19                   \\
+\multirow{3}{*}{ML\_MGSA\_Prob (Log Norm)}      & Linear                & 0.05                           & 114.72                 & 711.44              & 3.19                   \\
                                                                                            & Poly                  & 0.04                           & 114.79                 & 854.82              & 2.66                   \\
@@
 Compared to MGSA, the other algorithms reduce significantly the number of iterations in conditions when HV-indices are similar.
-However, the iteration number when using the models with probabilities (see last 3 rows of Table \ref{tab:1} at section ``Experiments for the case of fixed weights'') is greater than for the algorithm based on the separating hyperplane.
+However, the iteration number when using the models with probabilities (see rows ``ML\_MGSA\_Prob (Fixed weights)'' of Table \ref{tab:1}) is greater than for the algorithm based on the separating hyperplane.
 

### 2024-06-29 2024/OPTIMA/OPTIMA_2024_MCO.tex
OPTIMA add 4c alpha=0

@@
 \hline
-\textbf{}                                                                                  & \textbf{Model Kernel} & \textbf{$\alpha$} & \textbf{Avg.
-HV-index} & \textbf{Avg.
-iter.} & \textbf{Speedup iter.} \\ \hline
+\textbf{}                                                                  & \textbf{Model Kernel} & \textbf{$\alpha$} & \textbf{Avg. HV-index} & \textbf{Avg. iter.} & \textbf{Speedup iter.} \\ \hline
 ML\_MGSA                                                                   & -          & -    & 114.91  & 2269.82   & 1.00           \\ \hline
@@
 \multicolumn{6}{c}{\textbf{Experiments with 4-criterial MCO problems}}                                                                \\ \hline
-ML\_MGSA\_Dist                                                                             & Linear                & 0.01                           & 7751.43                & 1925.6              & 1.00                   \\
-\multirow{3}{*}{ML\_MGSA\_Prob (Log Norm)}      & Linear                & 0.08                           & 7729.71                & 1760.5              & 1.09                   \\
-                                                                                           & Poly                  & 0.08                           & 7796.2                 & 4121.7              & 0.47                   \\
-                                                                                           & RBF                   & 0.08                           & 7786.48                & 1364.6              & \textbf{1.41}          \\ \hline
+ML\_MGSA                                                                   & -          & -    & 7819.86 & 6046.4    & 1.00           \\ \hline
+ML\_MGSA\_Dist                                                             & Linear     & 0.01 & 7751.43 & 1925.6    & 3.14           \\
+\multirow{3}{*}{ML\_MGSA\_Prob (Log Norm)}                                 & Linear     & 0.08 & 7729.71 & 1760.5    & 3.43           \\
+                                                                           & Poly       & 0.08 & 7796.20 & 4121.7    & 1.47           \\
+                                                                           & RBF        & 0.08 & 7786.48 & 1364.6    & \textbf{4.43}  \\ \hline
 \end{tabular}%

### 2024-06-30 2024/OPTIMA/OPTIMA_2024_MCO.tex
last section

@@
 
-In the first experiment 100 bi-criterial MCO problems have been solved. As the partial criteria, the functions \cite{ML_MCO_2023} 
+In the first experiment 100 bi-criterial MCO problems have been solved. As the partial criteria, the functions $f(y)$ from \cite{ML_MCO_2023} were taken,
 \begin{equation}
@@
 \end{equation}
-were taken  where $0 \leq y_1, y_2 \leq 1$, parameters $-1 \leq A_{ij},B_{ij},C_{ij},D_{ij} \leq 1$ are independent uniformly distributed random values.
+where $0 \leq y_1, y_2 \leq 1$ and parameters $-1 \leq A_{ij},B_{ij},C_{ij},D_{ij} \leq 1$ are independent uniformly distributed random values.
 
-Computational experiments for techniques of Subsection \ref{ssec:42} which utilize probabilities of belonging to the Pareto set can be divided into 3 groups.
+Computational experiments for proposed techniques which utilize probabilities of belonging to the Pareto set can be divided into 3 groups.
 \begin{enumerate}
-	\item 	Experiments with the algorithm ML\_MGSA\_Prob (fixed weights).
-	\item 	Solving MCO problems by the method ML\_MGSA\_Prob (Adjustive weights));
+	\item 	Experiments with the algorithm ML\_MGSA\_Prob (fixed weights);
+	\item 	Solving MCO problems by the method ML\_MGSA\_Prob (Adjustive weights);
 	\item 	Experiments with changing weights in the course of ML\_MGSA\_Prob running and transformed probabilities (\ref{eq:18})  (ML\_MGSA\_Prob (Norm Log)).
@@
 In each series of experiments for these optimization algorithms, the coefficients $\alpha$ from (\ref{eq:16}) changed in the range $[0.1, 0.9]$.
-Comparison with the algorithm MGSA that does not use machine learning methods (in the characteristic (\ref{eq:16}) $\alpha = 0$) and with the algorithm ML\_MGSA\_Dist taking into account the distance to the separating hyperplane was carried out.
-The comparison was made on the base of  the calculated HV index characterizing quality of the obtained Pareto set approximation \cite{ML_MCO_2023,Evtushenko2014}.
+Comparison with the algorithm MGSA that does not use machine learning methods (i.e.
+with $\alpha = 0$ in the characteristic (\ref{eq:16})) and with the algorithm ML\_MGSA\_Dist taking into account the distance to the separating hyperplane was carried out.
+The comparison was made on the base of  the HV index characterizing quality of the obtained Pareto set approximation \cite{ML_MCO_2023,Evtushenko2014}.
 Further subsections include the results of only those calculations in which the $\alpha$ parameter provided the best quality of the solution found.
@@
 \hline
-\textbf{}                                                                  & \textbf{Model Kernel} & \textbf{$\alpha$} & \textbf{Avg. HV-index} & \textbf{Avg. iter.} & \textbf{Speedup iter.} \\ \hline
+\textbf{}                                                                  & \textbf{Model Kernel} & \textbf{$\alpha$} & \textbf{Avg. HV-index} & \textbf{Avg. iter.} & \textbf{Speedup} \\ \hline
 ML\_MGSA                                                                   & -          & -    & 114.91  & 2269.82   & 1.00           \\ \hline
@@
 Acceleration of the algorithm (in terms of reducing the number of search trials required to evaluate the Pareto set) is achieved through the application of machine learning methods.
-Namely, when using the SVM method, trial points in the criteria space are classified (each point is/is not an effective solution), on the basis of which a numerical estimate of the probability of belonging to the Pareto set is built.
+Namely, when using the SVM method, trial points in the criteria space are classified (each point is/is not an efficient solution), on the basis of which a numerical estimate of the probability of belonging to the Pareto set is built.
 Polynomial and radial basis function kernels in SVM allows one to adequately approximate Pareto front of a complex structure.
 
-Solving several series of MCO problems with complex multiextremal criteria has demonstrated that the proposed approach to evaluating the set of effective solutions is promising.
+Solving several series of MCO problems with complex multiextremal criteria has demonstrated that the proposed approach to evaluating the set of efficient solutions is promising.
 Further research aims to develop the proposed approach by consideration of other machine learning models (in particular, artificial neural networks) to better classify trials and approximate the Pareto set.

### 2024-06-30 2024/OPTIMA/OPTIMA_2024_MCO.tex
before submission

@@
 where $d_i$ is the distance from the $i$-th point to the hyperplane, $d_{max}$, $d_{min}$ are maximum and minimum distances to the hyperplane.
-
-The value $R_{ps}(i)$ from (\ref{eq:16}) for the interval $(x_{i-1},x_{i})$ is calculated as the sum of normalized distances (\ref{eq:17}) to points $x_i$ and $x_{i-1}$.
+The value $R_{ps}(i)$ from (\ref{eq:16}) for the interval $(x_{i-1},x_{i})$ is calculated as the sum of normalized distances (\ref{eq:17}) to $i$-th and $(i-1)$-th points.
 Let's denote this version of GSA as ML\_MGSA\_Dist.
@@
 Note that the classifier models of most machine learning libraries can return not only the class label, but also the probability of relevance to the class.
-Unlike the approach described in Subsection \ref{ssec:41} the characteristic $R_{ps}(i)$ is computed as the average of probabilities of situations when the points $x_i$ and $x_{i-1}$ from interval $(x_{i-1},x_{i})$ belong to the Pareto set.
+Unlike the approach described in Subsection \ref{ssec:41} the characteristic $R_{ps}(i)$ is computed as the average of probabilities of situations when the points
+%$y_i=y(x_i)$ and $y_{i-1}=y(x_{i-1})$
+correspond to boudaries of the interval $(x_{i-1},x_{i})$ belong to the Pareto set.
 
@@
 where $p_i$ is a probability of the event that the point $x_i$ belongs to the Pareto set, $p_{min}$, $p_{max}$ are minimum and maximum values of such probabilities.
-After transformation the function $R_{ps}(i)$ is calculated, as earlier, as the average value of probabilities of events in which the points $x_i$ and $x_{i-1}$ from interval $(x_{i-1}, x_i)$ belong to the Pareto set.
+After transformation the function $R_{ps}(i)$ is calculated, as earlier, as the average value of probabilities of events in which the points correspond to bounaries of the interval $(x_{i-1}, x_i)$ belong to the Pareto set.
 
@@
 
-Computations were conducted on the supercomputer ``Lobachevsky'' of the Nizhni Novgorod State University (operating system CentOS 7, control system SLURM).
-To obtain the executable program code the compilers Intel C++ 17.0.0 and python 3.9 were used.
-Computations were performed with the support of the package scikit-learn 0.24.2 and the optimization software iOpt \cite{iOpt_url}.
+Computations were conducted on the ``Lobachevsky'' supercomputer node (two Intel Sandy Bridge E5-2660 2.2 GHz processors and 64 Gb RAM) using the scikit-learn 0.24.2 library.
+The proposed algorithms were implemented in Python in intelligent optimization framework iOpt \cite{iOpt_url}.
+
+%Computations were performed with the support of the scikit-learn 0.24.2 framework.
+
+%of the Nizhni Novgorod State University (operating system CentOS 7, control system SLURM).To obtain the executable program code the compilers Intel C++ 17.0.0 and python 3.9 were used.
+and the optimization software iOpt \cite{iOpt_url}.
 

### 2024-06-30 2024/OPTIMA/OPTIMA_2024_MCO.tex
fix

@@
 \institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
-\email{a230172@unn.ru}, \email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, \email{vagris@unn.ru}}
+\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, \email{vagris@unn.ru}}
 %
@@
 \includegraphics[width=0.8\textwidth]{fig1.png}
-\caption{Heatmap of the function $R_ps(i)$ in the criteria space for algorithms with different machine learning models.} 
+\caption{Heatmap of the function $R_{ps}(i)$ in the criteria space for algorithms with different machine learning models.} 
 \label{fig:1}

### 2024-07-05 2024/OPTIMA/OPTIMA_2024_MCO.tex
OPTIMA fix bug
fix some mistakes

@@
 Evgeny Kozinov \Letter\orcidID{0000-0001-6776-0096} \and
-Konstantin Barkalov \Letter\orcidID{0000-0001-5273-2471} \and
+Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
 Vladimir Grishagin\orcidID{0000-0002-2884-3670}}
@@
 The best value of the partial criterion $f_i(y)$, $1 \leq i \leq s$, is supposed to correspond to the global minimum of this criterion in the domain $D$.
-
-We will denote this problem in the form
+Let's write this problem in the form
 \begin{equation}
 \label{eq:02a}
-  \Phi(y) = (f_1 (y),f_2 (y), \dots, f_s(y)) \to \min, y \in D
+  \Phi(y) = (f_1 (y),f_2 (y), \dots, f_s(y)) \to \min, \; y \in D.
 \end{equation}
@@
 \end{equation}
-where Lipschitz constants $L_i>0$, $1 \leq i \leq s$, are unknown \textit{a priopi}.
+where Lipschitz constants $L_i>0$, $1 \leq i \leq s$, are unknown a priopi.
 
@@
 
-For solving the optimization problem (\ref{eq:02a}), in our study the computational scheme \cite{ML_MCO_2023} has been applied. It consists of the following main stages.
+For solving the optimization problem (\ref{eq:02a}), in our study the computational scheme from \cite{ML_MCO_2023} has been applied. It consists of the following main stages.
  \begin{enumerate}
@@
 \label{eq:08}
-|\varphi(y(x')) - \varphi(y(x''))| \leq H \|x' - x''\|^{1/N} , x \in [0,1],
+|\varphi(y(x')) - \varphi(y(x''))| \leq H |x' - x''|^{1/N} , x \in [0,1],
 \end{equation}
-with the H{\" o}lder constant $H=2L\sqrt{N+3}$, where $L$ is the Lipschitz constant of $F(\lambda,y)$ in the domain $D$.
+with the H{\" o}lder constant $H=2L\sqrt{N+3}$ where $L$ is the Lipschitz constant of $F(\lambda,y)$ in the domain $D$.
 
@@
     \label{eq:09}
-    0 = x_0 < x_1 < \dots < x_i < \dots < x_{k-1} = 1.
+    0 = x_0 < x_1 < \dots < x_i < \dots < x_{k-1} = 1
 \end{equation}
@@
     \label{eq:11}
-    R(i) = \varrho_i + \frac{(z_i-z_{i-1})^2}{m^2 \varrho_i} - \frac{2 (z_i+z_{i-1})}{m}, 1 \leq i \leq k-1.
+    R(i) = \varrho_i + \frac{(z_i-z_{i-1})^2}{m^2 \varrho_i} - \frac{2 (z_i+z_{i-1})}{m}, 1 \leq i \leq k-1,
 \end{equation}
@@
 
-\textit{Rule 4.} Choose the subinterval $(x_{t-1} ,x_t)$ with maximal characteristic $R(t)$
+\textit{Rule 4.} Choose the subinterval $(x_{t-1} ,x_t)$ with maximum characteristic $R(t)$
 \begin{equation}
@@
     \label{eq:13}
-    x^{k} = \frac{x_t + x_{t-1}}{2} - sign(z_t - z_{t-1}) \frac{1}{2r} \left(\frac{|z_t - z_{t-1}|}{m} \right)^N,
+    x^{k} = \frac{x_t + x_{t-1}}{2} - sign(z_t - z_{t-1}) \frac{1}{2r} \left(\frac{|z_t - z_{t-1}|}{m} \right)^N
 \end{equation}
@@
     \label{eq:14}
-    \varrho < \varepsilon.
+    \varrho_t < \varepsilon
 \end{equation}
@@
     \label{eq:15}
-    \Omega_K=\{(x^q,y^q,f^q_1,\dots,f^q_s)^T: 0 \leq q \leq K\}.
+    \Omega_K=\{(x^q,y^q,f^q_1,\dots,f^q_s)^T: 0 \leq q \leq K\}
 \end{equation}
@@
     \label{eq:16}
-    R(i) = R_{gsa} (i) +  \alpha R_{ps} (i).
+    R(i) = R_{gsa} (i) +  \alpha R_{ps} (i)
 \end{equation}
-instead of the characteristic (\ref{eq:11}) where $R_{gsa}(i)$ coincides with the expression (\ref{eq:11}) and the additional term $R_{ps}(i)$ depends on closeness of the points $y(x_{i-1})$, $y(x_{i})$ to the Pareto set, namely, the closer these points to the Pareto set, the greater the value $R_{ps}(i)$.
+instead of the characteristic (\ref{eq:11}), where $R_{gsa}(i)$ coincides with the expression (\ref{eq:11}) and the additional term $R_{ps}(i)$ depends on closeness of the points $y(x_{i-1})$, $y(x_{i})$ to the Pareto set, namely, the closer these points to the Pareto set, the greater the value $R_{ps}(i)$.
 
@@
 \ref{fig:1} shows the behavior of the characteristic $R_{ps}(i)$ in the criteria space for different models.
-Red points are related to the Pareto set, orange those to the class  where points are not effective ones.
+Red points are related to the Pareto set, orange ones belong to the class of inefficient points.
 
@@
 Computations were conducted on the ``Lobachevsky'' supercomputer node (two Intel Sandy Bridge E5-2660 2.2 GHz processors and 64 Gb RAM) using the scikit-learn 0.24.2 library.
-The proposed algorithms were implemented in Python in intelligent optimization framework iOpt \cite{iOpt_url}.
+The proposed algorithms were implemented in Python in the \textit{i}ntelligent \textit{opt}imization framework iOpt \cite{iOpt_url}.
 

### 2024-07-07 2024/OPTIMA/OPTIMA_2024_MCO.tex
before submission

@@
 		\begin{matrix}
-		  f(y)= -(AB + CD)^{1/2} \\
-			AB =(\sum_{i=1}^7{\sum_{j=1}^7{[A_{ij} a_{ij} (y_1,y_2) + B_{ij} b_{ij} (y_1,y_2)])}})^2 \\
-			CD =(\sum_{i=1}^7{\sum_{j=1}^7{[C_{ij} a_{ij} (y_1,y_2) - D_{ij} b_{ij} (y_1,y_2)])}})^2 \\
+		  f(y)= -(AB + CD)^{1/2}, \\
+			AB =(\sum_{i=1}^7{\sum_{j=1}^7{[A_{ij} a_{ij} (y_1,y_2) + B_{ij} b_{ij} (y_1,y_2)])}})^2, \\
+			CD =(\sum_{i=1}^7{\sum_{j=1}^7{[C_{ij} a_{ij} (y_1,y_2) - D_{ij} b_{ij} (y_1,y_2)])}})^2, \\
 			a_{ij} (y_1,y_2) = \sin(\pi i y_1) \sin(\pi j y_2), \\

### 2024-07-25 2024/ITMO paper/sn-article.tex
fix first reviewer comments

@@
 \begin{remark}
-The Lipshitz condition will not be satisfied for the reduced one-dimensional function $\varphi(y(x))$.
+The Lipschitz condition will not be satisfied for the reduced one-dimensional function $\varphi(y(x))$.
 However, many one-dimensional Lipschitz optimization algorithms can be generalized to the case of H{\"o}lder function minimization.
@@
 
+\textcolor[rgb]{0,0,1}{
+\begin{remark}
+In this work we use piecewise-linear approximation $y_m(x)$ to the true curve $y(x)$ (Fig.~\ref{fig:Peano}).
+It is simple from the point of view of implementation and use, but does not reflect all the properties of the Peano-Hilbert curve.
+More complex approximations fit the curve better.
+For example, in \cite{Strongin2000} a non-univalent approximation was also proposed.
+It reflects the following property of the Peano-Hilbert curve $y(x)$: a point in $D$ can have several (but not more than $2^N$) inverse images in $[0,1]$.
+Using non-univalent approximation allows one to reduce the number of objective function evaluations required to solve the problem with a given accuracy (see \cite{Sergeyev2022, Lera2024}).
+\end{remark}}
+
 \subsection{The problem of discrete parameters}\label{sec_discr}
@@
 \]
-which translates any point of the interval $[0,S]$ to the region $D$ (the notation $E(x)$ corresponds to the integer part of the number $x$) and define the function
+which maps any point of the interval $[0,S]$ to the region $D$ (the notation $E(x)$ corresponds to the integer part of the number $x$) and define the function
 \begin{equation}\label{f}
@@
 The iOpt framework can be used to develop specialized decision support systems for solving the tasks of selecting optimal parameters for complex models or methods.
-The application area of the framework is focused on the hyperparameter tuning tasks, since this class of tasks involves the problems of multiextremality of the optimization criterion and the lack of a formula-based description of the model under study (a ``black box'' type model).
+\textcolor[rgb]{0,0,1}{In particular, it is focused on solving the problems of tuning hyperparameters of machine learning methods.
+In such problems there is a lack of analytical description of the objective function (``black box'' type problems), and the function itself can be multi-extremal.}
 

### 2024-07-26 2024/ITMO paper/sn-article.tex
before final check

@@
 
-An essential feature of the iOpt framework is the use of Lipschitz optimisation methods for hyperparameter tuning.
+An essential feature of the iOpt framework is the use of Lipschitz optimization methods for hyperparameter tuning.
 These methods are deterministic, provide fast convergence, and can be adapted to various types of special problem formulations.
@@
 Therefore, the values $z_i = f(x_i)$ at these points will be considered undefined and will not be used in the algorithm.
-%, а значения индекса -- равным 0, т.е. $\nu(x_i) = 0$.
+\textcolor[rgb]{0,0,1}{For the same treatment of boundary points of the interval $[0,S]$ we will also consider the values $f(0)$ and $f(S)$ to be undefined.
+The function values at boundary points will not be used in the algorithm.}
 
@@
 \textcolor[rgb]{0,0,1}{In particular, it is focused on solving the problems of tuning hyperparameters of machine learning methods.
-In such problems there is a lack of analytical description of the objective function (``black box'' type problems), and the function itself can be multi-extremal.}
+In such problems there is a lack of analytical description of the objective function and the function itself can be multi-extremal.}
 
@@
     \State compute the estimate of the H\"older constant $\mu$ by (\ref{mu})
-    \For{$i=0$ to $k$}
+    \For{$i=2$ to $k$}
         \State compute the characteristic $R(i)$ by (\ref{R})
     \EndFor
-    \State determine the number $t$, for which $R(t) = \max \left\{ R(i), 1 \leq i \leq k \right\}$
+    \State determine the number $t$, for which $R(t) = \max \left\{ R(i), 1 < i \leq k \right\}$
     \State compute the new trial point $x^{k+1}$ by (\ref{xk1})
@@
 \begin{enumerate}[1.]
-\item At the initialization stage, the values $(x^i = i, \nu^i=0, z^i)$, $ i, 0\leq i \leq S$, i.e., the boundary points of the intervals $(x_{i-1},x_i), 1<i\leq S$, are entered into the set $\Omega_k$ containing the search  information.
+\item At the initialization stage, the values $(x^i = i, \nu^i=0, z^i)$, $ 0\leq i \leq S$, i.e., the boundary points of the intervals $(x_{i-1},x_i), 1<i\leq S$, are entered into the set $\Omega_k$ containing the search  information.
+\textcolor[rgb]{0,0,1}{Recall that the values $z^i$ at integer points $x^i=i, 0\leq i \leq S,$ are supposed to be undefined and are not used later in the algorithm.}
 After that, trials are carried out at arbitrary internal points (e.g., midpoints) of the intervals $(x_{i-1},x_i), 1<i\leq S$; the trial results are also added to the search information.
@@
 \begin{proof}
-This statement is proved in a similar way to the theorem on the convergence of the index method of global optimisation from \citet{Strongin2000}, taking into account the relationship (\ref{f}) of the functions $f(x)$ and $\varphi(y,u_s)$, as well as the rules (\ref{index}), specifying the indices of the trial points.
+This statement is proved in a similar way to the theorem on the convergence of the index method of global optimization from \citet{Strongin2000}, taking into account the relationship (\ref{f}) of the functions $f(x)$ and $\varphi(y,u_s)$, as well as the rules (\ref{index}), specifying the indices of the trial points.
 \end{proof}

### 2024-08-03 2024/ITMO paper/sn-article.tex
Rev1 submitted

@@
 
-\textcolor[rgb]{0,0,1}{
 \begin{remark}
-In this work we use piecewise-linear approximation $y_m(x)$ to the true curve $y(x)$ (Fig.~\ref{fig:Peano}).
-It is simple from the point of view of implementation and use, but does not reflect all the properties of the Peano-Hilbert curve.
+In this work, we use piecewise-linear approximation $y_m(x)$ to the true curve $y(x)$ (Fig.~\ref{fig:Peano}).
+This approximation is simple from the point of view of implementation and use, but it does not reflect all the properties of the Peano-Hilbert curve.
 More complex approximations fit the curve better.
@@
 Using non-univalent approximation allows one to reduce the number of objective function evaluations required to solve the problem with a given accuracy (see \cite{Sergeyev2022, Lera2024}).
-\end{remark}}
+\end{remark}
 
@@
 Therefore, the values $z_i = f(x_i)$ at these points will be considered undefined and will not be used in the algorithm.
-\textcolor[rgb]{0,0,1}{For the same treatment of boundary points of the interval $[0,S]$ we will also consider the values $f(0)$ and $f(S)$ to be undefined.
-The function values at boundary points will not be used in the algorithm.}
+For treating the boundary points of the interval $[0,S]$ in a uniform manner, we will also consider the values $f(0)$ and $f(S)$ to be undefined.
+The function values at boundary points will not be used in the algorithm.
 
@@
 The iOpt framework can be used to develop specialized decision support systems for solving the tasks of selecting optimal parameters for complex models or methods.
-\textcolor[rgb]{0,0,1}{In particular, it is focused on solving the problems of tuning hyperparameters of machine learning methods.
-In such problems there is a lack of analytical description of the objective function and the function itself can be multi-extremal.}
+In particular, it is focused on solving the problems of tuning hyperparameters of machine learning methods.
+In such problems, there is a lack of analytical description of the objective function and the function itself may be multi-extremal.
 
@@
 \item At the initialization stage, the values $(x^i = i, \nu^i=0, z^i)$, $ 0\leq i \leq S$, i.e., the boundary points of the intervals $(x_{i-1},x_i), 1<i\leq S$, are entered into the set $\Omega_k$ containing the search  information.
-\textcolor[rgb]{0,0,1}{Recall that the values $z^i$ at integer points $x^i=i, 0\leq i \leq S,$ are supposed to be undefined and are not used later in the algorithm.}
+Recall that the values $z^i$ at integer points $x^i=i, 0\leq i \leq S,$ are assumed to be undefined and are not used later in the algorithm.
 After that, trials are carried out at arbitrary internal points (e.g., midpoints) of the intervals $(x_{i-1},x_i), 1<i\leq S$; the trial results are also added to the search information.

### 2024-08-03 2024/OPTIMA/OPTIMA_2024_MCO.tex

@@
 The complexity of these problems increases significantly in the case when the criteria are multiextremal.
-The paper presents several variations of the global optimization algorithm intended for solving ``black-box'' multiobjective problems with multiextremal criteria.
-The proposed methods are based on a combination of the information-statistical approach and machine learning procedures aimed at increasing the efficiency of constructing the Pareto set.
+The paper presents several variations of the global optimization algorithm intended for approximate solving ``black-box'' multiobjective problems with multiextremal criteria.
+The proposed methods are based on a combination of the information-statistical approach and machine learning procedures aimed at increasing the efficiency of constructing the Pareto set approximation.
 As a novelty, the paper describes a new approach to using machine learning models within the algorithm of multicriteria optimization.
@@
 
-\keywords{Multicriteria Problems \and Global Optimization \and Machine Learning.}
+\keywords{Multicriteria Problems \and Global Optimization \and Machine Learning \and Peano Curves.}
 \end{abstract}
@@
 \end{equation}
-where scalar functions $f_i(y)$, $1 \leq i \leq s$, are considered as the partial efficiency criteria, $y=(y_1,y_2, \dots ,y_N)$ is the vector of independent parameters belonging to the feasible domain $D \in R^N$ that is the hyperparallelepiped
+where scalar functions $f_i(y)$, $1 \leq i \leq s$, are considered as the partial efficiency criteria, $y=(y_1,y_2, \dots ,y_N)$ is the vector of independent parameters belonging to the feasible domain $D \subset R^N$ that is the hyperparallelepiped
 \begin{equation}
@@
 \label{eq:05}
-F(\lambda, y) = \max_{1 \leq i \leq s} \left\{\lambda_i f_i (y)\right\}, y \in D, \sum_{i=1}^s {\lambda_i} = 1, 0 \leq \lambda_i \leq 1.
+F(\lambda, y) = \max_{1 \leq i \leq s} \left\{\lambda_i f_i (y)\right\}, y \in D, \sum_{i=1}^s {\lambda_i} = 1, 0 < \lambda_i \leq 1.
 \end{equation}
@@
 
-\textbf{Dimensionality reduction.} For solving the optimization problems (\ref{eq:04}) the dimensionality reduction \cite{Gergel2019_2,Gergel2018,GergelKozinov2020} on the base of Peano curves is applied.
-The Peano curve $y(x)$ maps the close interval $[0,1]$ of the real axis onto the parallelepiped (\ref{eq:02}) unambiguously and continuously.
+\textbf{Dimensionality reduction.} For solving the optimization problems (\ref{eq:04}) the dimensionality reduction on the base of Peano curves is applied \cite{Gergel2019_2,Gergel2018,GergelKozinov2020,Lera2024,LERA2021125660}.
+The Peano curve $y(x)$ maps the closed interval $[0,1]$ of the real axis onto the parallelepiped (\ref{eq:02}) unambiguously and continuously.
 It enables to replace the multidimensional problem (\ref{eq:04}) with solving the equivalent univariate problem
@@
 %$y_i=y(x_i)$ and $y_{i-1}=y(x_{i-1})$
-correspond to boudaries of the interval $(x_{i-1},x_{i})$ belong to the Pareto set.
+corresponding to boudaries of the interval $(x_{i-1},x_{i})$ belong to the Pareto set.
 
@@
 
-\begin{figure}
+\begin{figure}[t!]
 \centering
@@
 
-First, when changing the optimization problem from set (\ref{eq:07}), a new weight calculation scheme for each of two classes of trial points is proposed.
+First, when changing the optimization problem from set (\ref{eq:07}), a new weight calculation scheme for each of the two classes of trial points is proposed.
 In this scheme, the weights of each class are calculated inversely proportional to the ratio of the number of all points to the number of points belonging to the Pareto set.
@@
 \ref{fig:1}.
-\begin{figure}
+\begin{figure}[t!]
 \centering
@@
 
-\textbf{Experiments with fixed weights of two trial classes.} The first series of experiments is aimed at comparison the algorithm ML\_MGSA\_Prob (Fixed weights) with results of running for methods MGSA and ML\_MGSA\_Dist.
+\textbf{Experiments with fixed weights of two trial classes.} The first series of experiments is aimed at comparison of the algorithm ML\_MGSA\_Prob (Fixed weights) with results of running for methods MGSA and ML\_MGSA\_Dist.
 
@@
 % \usepackage{graphicx}
-\begin{table}[ht]
+\begin{table}[t]
 \centering
@@
 Fig. \ref{fig:3} shows the change of the average HV-index as a function of the number of trials.
-\begin{figure}
+\begin{figure}[t!]
 \centering

### 2024-08-09 2024/OPTIMA/OPTIMA_2024_MCO.tex
fix some references
correct one sentence
fix the references

@@
 \usepackage{hyperref}
-
+\usepackage{color}
 \usepackage[misc,geometry]{ifsym}
@@
 \label{eq:05}
-F(\lambda, y) = \max_{1 \leq i \leq s} \left\{\lambda_i f_i (y)\right\}, y \in D, \sum_{i=1}^s {\lambda_i} = 1, 0 < \lambda_i \leq 1.
+F(\lambda, y) = \max_{1 \leq i \leq s} \left\{\lambda_i f_i (y)\right\}, y \in D, \sum_{i=1}^s {\lambda_i} = 1, 0 < \lambda_i < 1.
 \end{equation}
@@
 Moreover, the computation of the convolution (\ref{eq:05}) does not require the new calculation of criteria values at these points, because these values are taken from the matrix (\ref{eq:15}).
-After that, one can go to the basic iteration Rules 1-5.
+After that, one can start optimization algorithm according to the Rules 1-5.
 Such the scheme reduces significantly the number of criteria computations when jointly solving the family of problems (\ref{eq:04}) with convolution (\ref{eq:05}) for different weight coefficients $\lambda$ \cite{ML_MCO_2023,Gergel2018,GergelKozinov2020}.
@@
 It should be noted that in classical combinations of machine learning and optimization methods, the latter are used as an auxiliary internal techniques that improve the machine learning functioning.
+\textcolor[rgb]{0,0,1}{Other popular approaches use machine learning for auto-configuration (tuning the hyper-parameters) of optimization algorithms (see, e.g., \cite{Bezerra2020,Lopez2016} and the references given therein)}.
 In our approach, on the contrary, machine learning methods are a part of the optimization algorithms enhancing their work.

### 2024-08-09 2024/OPTIMA/OPTIMA_2024_MCO.tex
fix all the comments except two

@@
 
+\textcolor[rgb]{0,0,1}{\textit{Remark 1.} The Lipschitz constant of the objective function is assumed a priori unknown; it is adaptively estimated during the search.
+The value $M$ from (\ref{eq:10}) can be used in Lipschitz constant estimation.\\
+\textit{Remark 2.} It follows from relationship between the Lipschitz and H{\" o}lder constants of the functions $\varphi(y)$ and $\varphi(y(x))$, as well as from the theorem of convergence of the global search algorithm (see \cite{Strongin2000}), that parameter $r$ from (\ref{eq:10}) should satisfy the necessary condition $r>2^{2-1/N}$.
+In the test experiments described in Section~\ref{sec:5} we used the value $r=3$, which ensured the correct solution of the entire series of problems.
+In applied optimization problems the value $r=3$ also ensures the solution in the most cases.
+}
+
 %Theoretical justification of this algorithm is provided in \cite{Strongin2000}.
@@
 with $\alpha = 0$ in the characteristic (\ref{eq:16})) and with the algorithm ML\_MGSA\_Dist taking into account the distance to the separating hyperplane was carried out.
-The comparison was made on the base of  the HV index characterizing quality of the obtained Pareto set approximation \cite{ML_MCO_2023,Evtushenko2014}.
+The comparison was made on the base of the HV index characterizing quality of the obtained Pareto set approximation.
+\textcolor[rgb]{0,0,1}{The HyperVolume (HV) index corresponds to the volume of the region dominated by the Pareto set approximation and bounded by a given reference point \cite{ML_MCO_2023,Evtushenko2014}.
+The larger the hypervolume index, the better is the corresponding approximation.
+}
+
 Further subsections include the results of only those calculations in which the $\alpha$ parameter provided the best quality of the solution found.
@@
 
-Acceleration of the algorithm (in terms of reducing the number of search trials required to evaluate the Pareto set) is achieved through the application of machine learning methods.
+Acceleration of the algorithm (in terms of reducing the number of search trials required to evaluate \textcolor[rgb]{0,0,1}{the Pareto set approximation with a certain precision}) is achieved through the application of machine learning methods.
 Namely, when using the SVM method, trial points in the criteria space are classified (each point is/is not an efficient solution), on the basis of which a numerical estimate of the probability of belonging to the Pareto set is built.

### 2024-08-12 2024/OPTIMA/OPTIMA_2024_MCO.tex
add A.G. suggesions

@@
 As a novelty, the paper describes a new approach to using machine learning models within the algorithm of multicriteria optimization.
-The effectiveness of different versions of the proposed algorithm is estimated on the base of a representative computational experiment.
+Based on a representative computational experiment, the effectiveness of different versions of the proposed algorithm is estimated.
 
@@
 
-As it was mentioned above, the partial criteria contradictoriness leads to the necessity to introduce the set of non-dominated points called Pareto set as the general solution of the multicriteria problem \cite{Miettinen1999,Ehrgott2005,Pardalos2017}.
+As it was mentioned above, the contradictory nature of the subcriteria makes it necessary to introduce the set of non-dominated points called Pareto set as the general solution of the multicriteria problem \cite{Miettinen1999,Ehrgott2005,Pardalos2017}.
 Hereinafter, any point of the Pareto set will be called a \textit{partial solution} or just an \textit{efficient point}.
@@
 
-One of the ways how this can be done is connected with the modification of the starting iterations of GSA.
-Namely, if $P \geq 1$ problems (\ref{eq:07}) have been already solved and the matrix (\ref{eq:15}) has been formed, then as initial trial coordinates, the points $x^q$, $0 \leq q \leq K$, can be used with values $z^q = F(\lambda,y^q)$.
-Moreover, the computation of the convolution (\ref{eq:05}) does not require the new calculation of criteria values at these points, because these values are taken from the matrix (\ref{eq:15}).
-After that, one can start optimization algorithm according to the Rules 1-5.
-Such the scheme reduces significantly the number of criteria computations when jointly solving the family of problems (\ref{eq:04}) with convolution (\ref{eq:05}) for different weight coefficients $\lambda$ \cite{ML_MCO_2023,Gergel2018,GergelKozinov2020}.
+One way of achieving this is to modify the starting iterations of GSA.
+Specifically, if $P \geq 1$ problems (\ref{eq:07}) have already been solved and the matrix (\ref{eq:15}) has been formed, then the points $x^q$, $0 \leq q \leq K$, with values $z^q = F(\lambda,y^q)$ can be used as initial trial coordinates.
+Moreover, the computation of the convolution (\ref{eq:05}) does not require the recalculation of criteria values at these points, since these values are taken from the matrix (\ref{eq:15}).
+The optimization algorithm can then be started according to the Rules 1-5.
+This scheme significantly reduces the number of criteria computations when solving simultaneously the family of problems (\ref{eq:04}) with convolution (\ref{eq:05}) for different weight coefficients $\lambda$ \cite{ML_MCO_2023,Gergel2018,GergelKozinov2020}.
 
-The second approach also aims to further decreasing the number of trials required to find the Pareto set.
+The second approach also aims to further reduce the number of trials required to find the Pareto set.
 It suggests to change the form of the characteristics (\ref{eq:11}) in the GSA scheme.
@@
 
-After the construction of the separating hyperplane, for each trial point the $d_i$ distance is calculated and normalized according to the following expression:
+After constructing the separating hyperplane, for each trial point the $d_i$ distance is calculated and normalized according to the following expression:
 \begin{equation}
@@
 
-Solving several series of MCO problems with complex multiextremal criteria has demonstrated that the proposed approach to evaluating the set of efficient solutions is promising.
-Further research aims to develop the proposed approach by consideration of other machine learning models (in particular, artificial neural networks) to better classify trials and approximate the Pareto set.
+Solving several series of MCO problems with complex multiextremal criteria has demonstrated the potential of the proposed approach to evaluate the set of efficient solutions.
+Further research aims at extending the proposed approach by considering other machine learning models (in particular, artificial neural networks) to better classify the trials and approximate the Pareto set.
 

### 2024-08-25 2024/OPTIMA/OPTIMA_2024_MCO.tex
fix last comments

@@
 \title{Comparative efficiency of machine learning models for enhancing algorithms in solving multiextremal multicriteria problems\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.
-FSWR-2023-0034, and by by the Research and Education Mathematics Centre ``Mathematics for Future Technologies'', project no.
+FSWR-2023-0034, and by the Research and Education Mathematics Center ``Mathematics for Future Technologies'', project no.
 075-02-2024-1439.} }
@@
 
-\textcolor[rgb]{0,0,1}{\textit{Remark 1.} The Lipschitz constant of the objective function is assumed a priori unknown; it is adaptively estimated during the search.
-The value $M$ from (\ref{eq:10}) can be used in Lipschitz constant estimation.\\
+\textit{Remark 1.} The Lipschitz constant of the objective function is assumed a priori unknown; it is adaptively estimated during the search.
+The value $M$ from (\ref{eq:10}) can be used in Lipschitz constant estimation.
+
 \textit{Remark 2.} It follows from relationship between the Lipschitz and H{\" o}lder constants of the functions $\varphi(y)$ and $\varphi(y(x))$, as well as from the theorem of convergence of the global search algorithm (see \cite{Strongin2000}), that parameter $r$ from (\ref{eq:10}) should satisfy the necessary condition $r>2^{2-1/N}$.
@@
 In applied optimization problems the value $r=3$ also ensures the solution in the most cases.
-}
 
@@
 It should be noted that in classical combinations of machine learning and optimization methods, the latter are used as an auxiliary internal techniques that improve the machine learning functioning.
-\textcolor[rgb]{0,0,1}{Other popular approaches use machine learning for auto-configuration (tuning the hyper-parameters) of optimization algorithms (see, e.g., \cite{Bezerra2020,Lopez2016} and the references given therein)}.
+Other popular approaches use machine learning for auto-configuration (tuning the hyperparameters) of optimization algorithms (see, e.g., \cite{Bezerra2020,Lopez2016} and the references given therein).
 In our approach, on the contrary, machine learning methods are a part of the optimization algorithms enhancing their work.
@@
 The comparison was made on the base of the HV index characterizing quality of the obtained Pareto set approximation.
-\textcolor[rgb]{0,0,1}{The HyperVolume (HV) index corresponds to the volume of the region dominated by the Pareto set approximation and bounded by a given reference point \cite{ML_MCO_2023,Evtushenko2014}.
-The larger the hypervolume index, the better is the corresponding approximation.
-}
+The HyperVolume (HV) index is a traditional indicator of quality when approximating a Pareto set \cite{ML_MCO_2023,Evtushenko2014}.
+It corresponds to the volume of the region dominated by the Pareto set approximation obtained in the course of numerical solving.
+The greater the HV index, the better the approximation.
 
@@
 
-The following parameters are indicated in Table \ref{tab:1}: the kernel of the SVM model (column 2), parameter $\alpha$ (column 3), average values of the HV-index (column 4), the average number of trials spent by the algorithm (column 5), and the coefficient of decreasing the number of trials compared to ML\_MGSA (column 6).
+The following parameters are indicated in Table \ref{tab:1}: the kernel of the SVM model (column 2), parameter $\alpha$ (column 3), average values of the HV index (column 4), the average number of trials spent by the algorithm (column 5), and the coefficient of decreasing the number of trials compared to ML\_MGSA (column 6).
 The factors $r$ and $\varepsilon$ from (\ref{eq:10}) and (\ref{eq:14}) were the same for all the algorithms.
@@
 \hline
-\textbf{}                                                                  & \textbf{Model Kernel} & \textbf{$\alpha$} & \textbf{Avg. HV-index} & \textbf{Avg. iter.} & \textbf{Speedup} \\ \hline
+\textbf{}                                                                  & \textbf{Model Kernel} & \textbf{$\alpha$} & \textbf{Avg. HV index} & \textbf{Avg. iter.} & \textbf{Speedup} \\ \hline
 ML\_MGSA                                                                   & -          & -    & 114.91  & 2269.82   & 1.00           \\ \hline
@@
 
-Fig. \ref{fig:3} shows the change of the average HV-index as a function of the number of trials.
 \begin{figure}[t!]
@@
 \includegraphics[width=0.8\textwidth]{fig3.png}
-\caption{Comparison of HV-indices for the case of fixed weights.} 
+\caption{Comparison of HV indices for the case of fixed weights.} 
 \label{fig:3}
@@
 
-Compared to MGSA, the other algorithms reduce significantly the number of iterations in conditions when HV-indices are similar.
+Fig. \ref{fig:3} shows the change of the average HV index as a function of the number of trials.
+Compared to MGSA, the other algorithms reduce significantly the number of iterations in conditions when HV indices are similar.
 However, the iteration number when using the models with probabilities (see rows ``ML\_MGSA\_Prob (Fixed weights)'' of Table \ref{tab:1}) is greater than for the algorithm based on the separating hyperplane.
@@
 \includegraphics[width=0.8\textwidth]{fig4.png}
-\caption{HV-indices comparison  in the case of adjusted weights} 
+\caption{HV indices comparison  in the case of adjusted weights} 
 \label{fig:4}
@@
 \includegraphics[width=0.8\textwidth]{fig5.png}
-\caption{Comparison of  HV-indices for methods with probabilities (\ref{eq:18})} 
+\caption{Comparison of  HV indices for methods with probabilities (\ref{eq:18})} 
 \label{fig:5}
@@
 
-The results of the experiments allow us to conclude that the method ML\_MGSA\_Prob (Log Norm) in this series of experiments can significantly reduce the number of trials with a similar value of HV-index compared to ML\_MGSA\_Dist.
+The results of the experiments allow us to conclude that the method ML\_MGSA\_Prob (Log Norm) in this series of experiments can significantly reduce the number of trials with a similar value of HV index compared to ML\_MGSA\_Dist.
 The best results were achieved using models with Linear and RBF kernels, while the latter model demonstrates better quality than the version of the algorithm with the technique \cite{ML_MCO_2023} based on distances to the hyperplane.
@@
 
-Acceleration of the algorithm (in terms of reducing the number of search trials required to evaluate \textcolor[rgb]{0,0,1}{the Pareto set approximation with a certain precision}) is achieved through the application of machine learning methods.
+Acceleration of the algorithm (in terms of reducing the number of search trials required to evaluate the Pareto set approximation with a certain precision) is achieved through the application of machine learning methods.
 Namely, when using the SVM method, trial points in the criteria space are classified (each point is/is not an efficient solution), on the basis of which a numerical estimate of the probability of belonging to the Pareto set is built.
@@
 Further research aims at extending the proposed approach by considering other machine learning models (in particular, artificial neural networks) to better classify the trials and approximate the Pareto set.
+Additionally, we intend to expand the basis of computational experiment by using new test sets, for example, the benchmarks \cite{Deb2003} and considering the statistical analysis to estimate test results more qualitative.
 

### 2024-08-26 2024/OPTIMA/OPTIMA_2024_MCO.tex
submitted

@@
 
-%\begin{equation}
-%\label{eq:06}
-%F(y) = \{ F(\lambda_1,y),F(\lambda_2,y), \dots ,F(\lambda_\tau,y)\}
-%\end{equation}
-
 In the case of partial criteria positivity the global minimizer of the convolution (\ref{eq:05}) is the Pareto-optimal point.
@@
 In the test experiments described in Section~\ref{sec:5} we used the value $r=3$, which ensured the correct solution of the entire series of problems.
-In applied optimization problems the value $r=3$ also ensures the solution in the most cases.
-
-%Theoretical justification of this algorithm is provided in \cite{Strongin2000}.
+In applied optimization problems the value $r=3$ also provides the correct solution in most cases.
 
@@
 Unlike the approach described in Subsection \ref{ssec:41} the characteristic $R_{ps}(i)$ is computed as the average of probabilities of situations when the points
-%$y_i=y(x_i)$ and $y_{i-1}=y(x_{i-1})$
-corresponding to boudaries of the interval $(x_{i-1},x_{i})$ belong to the Pareto set.
+corresponding to boundaries of the interval $(x_{i-1},x_{i})$ belong to the Pareto set.
 
 A significant advantage of this approach is the possibility to easily replace both the machine learning model itself and its learning rules.
-In the framework of the study, several variants of building models that potentially improve the factor $R_{ps}(i)$ to beter solve the MCO problems.
+In the framework of the study, several variants of building models that potentially improve the factor $R_{ps}(i)$ to better solve the MCO problems.
 
@@
 
-%Computations were performed with the support of the scikit-learn 0.24.2 framework.
-
-%of the Nizhni Novgorod State University (operating system CentOS 7, control system SLURM).To obtain the executable program code the compilers Intel C++ 17.0.0 and python 3.9 were used.
-and the optimization software iOpt \cite{iOpt_url}.
-
 In the first experiment 100 bi-criterial MCO problems have been solved. As the partial criteria, the functions $f(y)$ from \cite{ML_MCO_2023} were taken,
@@
 
-% Please add the following required packages to your document preamble:
-% \usepackage{multirow}
-% \usepackage{graphicx}
 \begin{table}[t]
@@
 %
-%\begin{thebibliography}{8}
-%\bibitem{ref_article1}
-%Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-%
-%\bibitem{ref_lncs1}
-%Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-%F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-%Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-%
-%\bibitem{ref_book1}
-%Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-%Location (1999)
-%
-%\bibitem{ref_proc1}
-%Author, A.-B.: Contribution title. In: 9th International Proceedings
-%on Proceedings, pp. 1--2. Publisher, Location (2010)
-%
-%\bibitem{ref_url1}
-%LNCS Homepage, \url{http://www.springer.com/lncs}, last accessed 2023/10/25
-%\end{thebibliography}
 \end{document}

### 2024-09-29 2024/NumMet/NMP_authorGuidelinesRu_2023.tex
add first version of convergence proof

@@
 %===========================  ШАПКА СТАТЬИ =================================
-\Article{Алгоритм липшицевой глобальной оптимизации\break
-    для частично определенной функции}
-    {Lipschitz global optimization algorithm \break
-    of a partially defined function}
+\Article{Сходимость алгоритма липшицевой глобальной\break
+    оптимизации для частично определенной функции}
+    {Convergence of the Lipschitz Global Optimization Algorithm\break
+    for a Partially Defined Function}
 
 \Abstract{В статье обсуждается проблема нахождения глобального минимума липшецевой функции, которая может быть частично определена в области поиска.
-Такое явление возможно в силу особенностей оптимизируемого объекта и метода моделирования (например, численная нестабильность метода моделирования).
+Такое явление возможно в силу особенностей оптимизируемого объекта и метода моделирования (например, численная нестабильность).
 В некоторых случаях эти подобласти известны, но в большинстве случаев информация о них отсутствует.
-В статье изложено краткое описание алгоритма глобального поиска для решения такого класса задач.
-Даны теоретические обоснования сходимости алгоритма.
-Проведены численные эксперименты, подтверждающие эффективность предложенного алгоритма.}
+В статье изложено описание модификации алгоритма глобального поиска для решения такого класса задач.
+Даны теоретические обоснования сходимости алгоритма.}
     {The paper discusses the problem of finding the global minimum of a Lipschitz function that may be partially defined in the search domain.
-This phenomenon is possible due to the nature of the optimized object and of the simulation method (for example, numerical instability of the simulation method).
-In some cases these subareas are known, but in most cases the information about them is missing.
-The paper provides a brief description of the global search algorithm for solving such class of problems.
+This phenomenon is possible due to the nature of the optimized object and of the simulation method (for example, numerical instability).
+In some cases, these subareas are known, but in most cases the information about them is missing.
+The paper provides a modification the global search algorithm for solving this class of problems.
 Theoretical justifications for the convergence of the algorithm are given.
-Numerical experiments confirming the efficiency of the proposed algorithm were carried out.}
+The results of numerical experiments are presented.}
 
@@
         многоэкстремальные функции, 
-        частично определенные функции.}
-        {Lipschitz global optimization, 
-        Multiextremal functions, 
-        Partially defined functions.}
+        частично определенные функции,
+        условия сходимости.}
+        {lipschitz global optimization, 
+        multiextremal functions, 
+        partially defined functions,
+        convergence conditions.}
 
@@
 \Citation{Усова~М.А., Баркалов~К.А.
-    Алгоритм липшицевой глобальной оптимизации
+    Сходимость алгоритма липшицевой глобальной оптимизации
     для частично определенной функции~//
@@
     {M.~A.~Usova, K.~A.~Barkalov
-    ``Lipschitz global optimization algorithm
-    of a partially defined function'',
+    ``Convergence of the Lipschitz Global Optimization Algorithm
+    for a Partially Defined Function'',
     Numerical Methods and Programming. \textbf{??} (?), 
@@
 [\ref{rfa:rulit:Grishagin2016_2},\ref{rfa:rulit:Jones2021},\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Birect2020},\ref{rfa:rulit:Sergeyev2017}].
-Проведенные сравнения показывают, что детерминированные алгоритмы превосходят (по разным критериям) широко распространенные nature-inspired algorithms [\ref{rfa:rulit:Liberti2005},\ref{rfa:rulit:Sergeyev2018}].
+Проведенные сравнения показывают, что детерминированные алгоритмы превосходят (по разным критериям) широко распространенные природно-естественные алгоритмы [\ref{rfa:rulit:Liberti2005},\ref{rfa:rulit:Sergeyev2018}].
 
 Данная работа продолжает развитие одного из эффективных детерминированных методов решения задач липшицевой глобальной оптимизации – информационно-статистического алгоритма глобального поиска [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}].
-Данный алгоритм предполагает решение многомерных задач посредством их сведения к задачам одномерной оптимизации при помощи space-filling curves (Peano curves).
+Данный алгоритм предполагает решение многомерных задач посредством их сведения к задачам одномерной оптимизации при помощи кривой заполняющей пространство (кривые Пеано).
 
@@
 
-Однако в последнее время актуальной становится принципиально новая проблема прикладных задач – численная нестабильность исследуемых мо-делей в некоторых (заранее не известных) подобластях области изменения параметров.
+Однако в последнее время актуальной становится принципиально новая проблема прикладных задач – численная нестабильность исследуемых моделей в некоторых (заранее не известных) подобластях области изменения параметров.
 В таких подобластях невозможно корректно провести численное моделирование и вычислить значение целевой функции задачи.
-Данное явление можно интерпретировать либо как наличие в задаче некоторых скрытых ограничений (hidden constraints) [\ref{rfa:rulit:Stripinis2021}], или как наличие неизвестных разрывных областей [\ref{rfa:rulit:Audet2022}], или как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Sergeyev2003},\ref{rfa:rulit:Strongin2020}].
+Данное явление можно интерпретировать или как наличие в задаче некоторых скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], или как наличие неизвестных разрывных областей [\ref{rfa:rulit:Audet2022}], или как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Sergeyev2003},\ref{rfa:rulit:Strongin2020}].
 В такой постановке задача оптимизации существенно усложняется, т.к.
@@
 
-В статье отражены результаты нового направления исследований, связан-ного с использованием импутации\footnote{Термин импутиция (imputation) в статье используется в соответствии с устоявшейся терминологией в области машинного обучения, связанной с восстановлением недостающих данных.} (восстановления) значений целевой функции в ходе работы алгоритма глобального поиска при решении задач с частично определенной целевой функцией.
-Предложенный подход основан на адаптации вычислительных правил алгоритма глобального поиска.
-Эффективность предложенного подхода демонстрируется численно.
+В статье приводится доказательство сходимости одного из предложенных ранее подходов к модификации алгоритма глобального поиска на случай частично определенной целевой функции [\ref{rfa:rulit:Usova2024}].
+Также продемонстрированы новые результаты численных экспериментов.
 
 \section{Постановка задачи}
-В общем виде задача глобальной оптимизации может быть сформулирова-на следующим образом:
+В общем виде задача глобальной оптимизации может быть сформулирована следующим образом:
 \begin{equation}\label{eq1} 
@@
 Используя кривые типа развертки Пеано, однозначно отображающие отрезок $[0,1]$ на $N$-мерный единичный гиперкуб
-$$
+\begin{equation}\label{eq2_} 
 D=\left\{ y \in R^N: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\},
-$$
+\end{equation}
 исходную задачу (\ref{eq1}) можно редуцировать к одномерной задаче
@@
 \end{equation}
-что позволяет применить для ее решения эффективные алгоритмы одномер-ной оптимизации.
+что позволяет применить для ее решения эффективные алгоритмы одномерной оптимизации. О целевой функции $\phi (y)$ мы делаем следующие предположения.
 
-О целевой функции задачи мы делаем следующие предположения.
-
-Во-первых, целевая функция может быть многоэкстремальной, недиффе-ренцируемой и, более того, заданной в форме ``черного ящика'' (т.е.
-в виде некоторой подпрограммы, на вход которой подается аргумент, а выходом является соответствующее значение функции).
-
-Во-вторых, каждое вычисление функции в некоторой точке допустимой области может требовать значительных вычислительных ресурсов.
-
-В-третьих, целевая функция удовлетворяет условию Липщица
+\begin{enumerate}
+\item{Целевая функция может быть многоэкстремальной, недифференцируемой и, более того, заданной в форме ``черного ящика'' (т.е.
+в виде некоторой подпрограммы, на вход которой подается аргумент, а выходом является соответствующее значение функции).}
+\item{Каждое вычисление функции в некоторой точке допустимой области может требовать значительных вычислительных ресурсов.}
+\item{Целевая функция удовлетворяет условию Липщица
 \begin{equation}\label{eq3} 
@@
 \end{equation}
-где $\rho(x',x'') =  |x' - x''|^{1/N}$ и $N$ -- размерность исходной многомерной задачи, а коэффициент $K$ связан с константой Липшица $L$ соотношением $K \leq 2L\sqrt {N+3}$ [\ref{rfa:rulit:Strongin2000}].
+где $\rho(x',x'') =  |x' - x''|^{1/N}$ -- метрика Гёльдера, $N$ -- размерность исходной многомерной задачи, а коэффициент $K$ связан с константой Липшица $L$ соотношением $K \leq 2L\sqrt {N+3}$ [\ref{rfa:rulit:Strongin2000}].}
+\item{В некоторых подобластях $I_1, I_2, ..., I_k: I_1 \cup I_2 \cup ...
+\cup I_k = I$ области поиска $D$ (в частном случае, в одной или нескольких ее точках) целевая функция может быть не определена.
+Тогда функция $\phi(y)$ определена и вычислима лишь в некоторой области $Q = D \backslash I$ положительного объёма.
+Отметим, что исходя из опыта решения прикладных оптимизационных задач [\ref{rfa:rulit:Barkalov2022},\ref{rfa:rulit:Gubaydullin2022}], суммарный объем области невычислимости $I$ составляет небольшую долю объема области поиска $D$.}
+\end{enumerate}
 
-Различные варианты алгоритмов для решения такого класса задач и соот-ветствующая теория сходимости представлены в работах [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}].
+Последнее предположение делает невозможными применение классического информационно-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска  [\ref{rfa:rulit:Strongin2000}] или других методов липшицевой оптимизации [\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Sergeyev2017}].
+На случай решения задач с невсюду вычислимой целевой функцией нами была разработана модификация алгоритма глобального поиска (АГП-Н) [\ref{rfa:rulit:Usova2024}].
+Далее будет приведено описание одной из версий модификации алгоритма и указаны условия сходимости данного метода.
 
-В-четвертых, в некоторых подобластях области поиска $D$ (в частном случае, в одной или нескольких ее точках) целевая функция может быть не определена.
+\section{Алгоритм решения задач с частично определенной целевой функцией}
 
-Исходя из опыта решения прикладных оптимизационных задач [\ref{rfa:rulit:Barkalov2022},\ref{rfa:rulit:Gubaydullin2022}], мы будем предполагать, что суммарный объем областей невычислимости составляет небольшую долю объема области поиска $D$ (около 10-20\%)\footnote{При тестировании алгоритма мы рассматривали и более сложные задачи, в которых объём невычислимой области достигал 50-80\%.Задачи такого рода также успешно решались предложенным алгоритмом, но для этого требовалось большое количество испытаний, поскольку метод пытался исследовать всю область невычислимости.
-В связи с этим в правила построения тестовых задач для размерности $N=5$ были внесены изменения таким образом, чтобы области невычислимости были не повсеместно по контуру, а только в угловых частях области поиска}.
+Для реализации модификации алгоритма рассматрвается простейший подход - игнорировать отсутствие значения целевой функции в точке.
+Расчёт характеристики интервала с двумя невычислимыми или с невычислимой и граничной точками производить по длине интервала с использованием понижающей константы.
+Расчёт характеристики интервала с невычислимой и вычислимой точками производить по правилам работы с граничным интервалом.
 
-Последнее предположение делает невозможными применение классического информационно-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска  [\ref{rfa:rulit:Strongin2000}] или других методов липшицевой оптимизации [\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Sergeyev2017}] и требует разработки модификации на случай решения задач с не всюду определенной целевой функцией.
+\begin{figure}
+\includegraphics[width=\textwidth]{pic/fig2.png}
+\caption{Примеры распределения поисковых испытаний в областях невычислимости} \label{fig2}
+\end{figure}
 
-\section{Метод решения задач с частично определенной целевой функцией}
+Приведем подробное описание модификации алгоритма глобального поиска для невсюду вычислимой целевой функции (АГП-Н) для решения задач (\ref{eq2}).
 
-\subsection{Алгоритм глобального поиска}
-
-%В рамках предлагаемого подхода для решения задач (\ref{eq2}) используется информационно-статистическая теория глобального поиска.
-Основная идея алгоритма в использовании накопленной информации для определения очередного интервала, на котором абсолютный минимум наиболее вероятен.
-Очередная точка в этом интервале соответствует математическому ожиданию положения минимума.
-Приведем более подробное описание \textbf{алгоритма глобального поиска} (АГП) для решения задач (\ref{eq2}).
+\subsection{Описание алгоритма АГП-Н}
 
@@
   \begin{cases}
-    -1 & {\quad \text{if } x_i \text{ -- граничная точка,}}\\
-    1  & {\quad \text{if } x_i \text{ -- внутренняя точка.}}
+    -1, & {\quad \text{if } x_i \text{ -- граничная точка}},\\
+    0, & {\quad \text{if } x_i \text{ -- невычислимая точка}},\\
+    1, & {\quad \text{if } x_i \text{ -- внутренняя точка}}.
   \end{cases}
 \end{equation}
-Точки $x_0=0$ и $x_{k}=1$ введены дополнительно (значения $z_0$ и $z_{k}$ не определены) для удобства последующих обозначений.
+Точки $x_0=0$ и $x_{k}=1$ введены дополнительно (значения $z_0$ и $z_{k}$ не определены).
 
@@
 \begin{equation}\label{eq7} 
-\mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i}, 2 \leq i \leq k \right\} , \Delta _i= (x_i-x_{i-1})^{1/N},
+\mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i},\quad v(x_{i-1})=v(x_i)=1, \quad 2 \leq i \leq k  \right\},\quad \Delta _i= (x_i-x_{i-1})^{1/N},
 \end{equation}
-для константы Гёльдера $K$ из (\ref{eq4}) для редуцированной функции $f(x)$, где $r>1$ параметр надёжности ; если $\mu=0$, то принять $\mu=1$.
+для константы Гёльдера $K$ из (\ref{eq4}) для редуцированной функции $f(x)$, где $r>1$ -- параметр надёжности. Если $\mu=0$, то принять $\mu=1$.
+
 \textit{Правило 3.} Определить текущее лучшее значение целевой функции
 \begin{equation}\label{eq8} 
-z^*=\min \left\{ f(x_i): 1\leq i \leq k \right\}.
+z^*= \min \left\{ f(x_i): v(x_i)=1, 0 < i < k \right\}.
 \end{equation}
@@
 \begin{equation}\label{eq9} 
-R(x_i)=
+R(i)=
   \begin{cases}
+    {(1-\frac{1}{r})}^2 \Delta _i,& {\quad v(x_i)=v(x_{i-1})=0 \text{ или } v(x_i)=0 \text{ и } v(x_{i-1})=-1}\\
+    {} & {\quad \text{или } v(x_i)=-1  \text{ и } v(x_{i-1})=0},\\
     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1},\\
-    2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1},\\
-    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
+    2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad  v(x_{i-1})=-1 \text{ или } v(x_{i-1})=0, v(x_i)=1},\\
+    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1 \text{ или } v(x_i)=0.}
   \end{cases}
@@
 \]
+Если максимальную характеристику имеют несколько интервалов, отдавать предпочтение вычислимым интервалам с минимальным номером.
 
-\textit{Правило 6.} Провести очередное испытание в серединной точке интервала $(x_{t-1},x_t)$, если индексы его концевых точек не совпадают, т.е.
+\textit{Правило 6.} Провести очередное испытание в серединной точке интервала $(x_{t-1},x_t)$, если индексы его концевых точек не совпадают или оба соответсвуют невычислимыми, т.е.
 \begin{equation}\label{eq10} 
-x^{k+1}=\frac {x_t+x_{t-1}}{2}, v(x_{t-1})\neq v(x_t).
+x^{k+1}=\frac {x_t+x_{t-1}}{2},\quad v(x_{t-1})\neq v(x_t) \text{ или } v(x_{t-1})=v(x_t)=0.
 \end{equation}
@@
 \begin{equation}\label{eq11} 
-x^{k+1}= \frac {x_t+x_{t-1}}{2} -  \text{sign} {(z_t-z_{t-1})} \frac{1}{2r} \left[\frac {{|z_t-z_{t-1}|}}{\mu} \right]^N, v(x_{t-1})=v(x_t).
+x^{k+1}= \frac {x_t+x_{t-1}}{2} -  \text{sign} {(z_t-z_{t-1})} \frac{1}{2r} \left[\frac {{|z_t-z_{t-1}|}}{\mu} \right]^N.
 \end{equation}
@@
 
-\subsection{Подходы к восстановлению значений целевой функции}
-Описанная ранее проблема не всюду определенной целевой функции является частичным аналогом уже давно существующей проблемы пропущенных значений в Data Science, причинами которой могут быть, например, технические проблемы или сборка датасета из нескольких источников с различными наборами параметров.
+\subsection{Условия сходимости и оценка эффективности}
 
-Стоит отметить, хотя некоторые алгоритмы машинного обучения умеют принимать во внимание и даже восстанавливать пропущенные значения в данных, например, LightGBM имеет режим игнорирования пропущенных значений, а XGBoost восстанавливает данные за счёт уменьшения функции потерь при обучении, большинство моделей не могут обрабатывать пропущенные значения и требуют валидные данные без $NaN$ или «missing» значений, что делает данную проблему достаточно серьёзной.
+Различные варианты алгоритмов для решения задач липшицевой оптимизации и соответствующая теория сходимости представлены в работах [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}].
+Приведём формулировку теоремы сходимости для АГП.
 
-Подходы к импутации (восстановлению) пропущенных значений в Data Science, предложенные специалистами в данной сфере, самые разные:
+\textbf{Теорема (о сходимости АГП)}.
+Пусть $\overline{y}=y(\overline{x})$ есть предельная точка последовательности испытаний $\{y^k\}=\{y(x^k)\}, \{x^k\} \subset [0,1]$, порождаемой АГП при минимизации липшицевой с константой $L$ функции $\phi(y), y = y(x) \in D$.
+Тогда если на некотором шаге поиска выполнено условие
+\[
+r\mu > 2^{3-1/N}L\sqrt{N+3},
+\]
+то любая точка глобального минимума $y^*$ является предельной точкой последовательности $\{y^k\}$ и любая точка $\overline{y}$ является точкой глобального минимума функции $\phi(y)$.
 
-\begin{itemize}[itemsep=0pt,parsep=2pt,topsep=2pt,partopsep=0pt]
-\item{удаление каждого наблюдения, содержащего одно или несколько пропущенных значений (крайнее средство, алгоритм теряет доступ к полезной информации, содержащейся в непропущенных значениях наблюдений);}
-\item{замена пропущенных значений средним/медианой (не работает с качественными переменными, не учитывает корреляцию данных);}
-\item{замена самым часто встречающимся значением или константой (отличается от использования медианы только работой с качественными параметрами);}
-\item{замена данных, используя метод $k$ ближайших соседей или kNN (в отличие от первых учитывает корреляцию между параметрами, чувствителен к выбросам, вычислительно дорого);}
-\item{множественная импутация данных (MICE) – импутация каждого значения проводится не один раз, а много (позволяет понять, насколько надёжно или ненадёжно предложенное значение);}
-\item{импутация данных с помощью глубокого обучения (восстанавливает недостающие значения за счет тренировки нейронной сети на тех точках, для которых есть все параметры, на больших данных вычислительно дорого).}
-\end{itemize}
-Общим недостатком является отсутствие универсального алгоритма, необходимость для каждой конкретной задачи искать наиболее подходящие методы.
+\textbf{Замечание}. Теорема соответствует случаю пустой области невычислимости целевой функции, т.е. $I = \emptyset$ и $Q = D$.
 
-Для импутации значений целевой функции может рассматриваться в целом схожий набор подходов, однако учитывающий липшицевость функции и использующий информационно-статистическую теорию, что позволяет подобрать универсальный алгоритм для данного класса задач.
-Далее точки, в которых обнаружена невычислимость, интерпретируются как имеющие нулевой индекс и (\ref{eq6}) заменяется расширенной формулой
+\textbf{Лемма (о попадании испытаний в область вычислимости)}.
+В последовательности испытаний, порождаемом алгоритмом АГП-Н при минимизации функции $\phi(y), y \in D$, всегда найдется такой номер итерации $h$, что в данной точке целевая функция будет определена и вычислима, т.е.
+$v(x^h) = 1$.
 
-\begin{equation}\label{eq12} 
-v_i=v(x_i)=
-  \begin{cases}
-    -1, & {\quad \text{if } x_i \text{ -- граничная точка}},\\
-    0, & {\quad \text{if } x_i \text{ -- невычислимая точка}},\\
-    1, & {\quad \text{if } x_i \text{ -- внутренняя точка}}.
-  \end{cases}
-\end{equation}
+\textbf{Доказательство}.
+Пусть $y' \in Q$ -- внутренняя точка множества $Q$.
+Значит существует такая точка $x' \in [0,1]$, являющаяся прообразом $y' = y(x')$, и такой отрезок $[\alpha, \beta] \subset [0, 1]$, содержащий точку $x'$, что для любой точки $x \in [\alpha, \beta]$ ее образ попадает в область вычислимости $y(x) \in Q$ (Рис.
+\ref{fig_1}).
 
-Опишем возможные подходы к реализации модификации.
-
-\textbf{Первый подход.} Игнорировать отсутствие значения целевой функции в точке, расчёт характеристики интервала хотя бы с одной невычислимой точкой производить по длине интервала, т.е.
-\begin{equation}\label{eq13} 
-R(i)=\Delta _i, \text{ if } v(x_i)=0 \text{ или } v(x_{i-1})=0.
-\end{equation}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{pic/fig1.png}
-\caption{Примеры распределения поисковых испытаний в областях невычислимости при первом подходе к импутации значений} \label{fig1}
+\begin{figure}[h]
+\includegraphics[width=200pt]{pic/fig_1.jpg}
+\caption{Существование вычислимого отрезка на кривой Пеано} \label{fig_1}
 \end{figure}
 
-Существенным минусом данного подхода является большое число избыточных испытаний после попадания в область невычислимости за счёт большого значения характеристики, получаемой по формуле (\ref{eq13}).
+Значит на каждом шаге поиска $k \geq 1$ существует некоторый интервал $(x_{j-1}, x_{j})$ такой, что $x' \in (x_{j-1}, x_{j})$.
+Если $y(x_{j-1})$ или $y(x_{j})$ принадлежат области $Q$, то $v(x_{j-1}) = 1$ или $v(x_j) = 1$ и всё доказано.
 
-\textbf{Второй подход.} Игнорировать отсутствие значения целевой функции в точке, расчёт характеристики интервала с двумя невычислимыми или с невычислимой и граничной точками производить по длине интервала, умноженной на понижающую константу, т.е.
-\begin{equation}\label{eq14} 
-R(i)={(1-{1}/{r})}^N \Delta _i,
-\end{equation}
-если $v(x_i)=v(x_{i-1})=0$ или $v(x_i)=0$ и $v(x_{i-1})=-1$ или $v(x_i)=-1$ и $v(x_{i-1})=0$.
+Иначе, $R(j)=(1-\frac{1}{r})^2\Delta_j>0$ и в этот интервал попадёт точка одного из следующих испытаний.
+При $k \to \infty$ алгоритм будет порождать последовательность интервалов, сходящихся к $x'$, а значит через некоторое количество итераций $h$ граница одного из интервалов попадёт в $[\alpha, \beta]$ и $v(x_{j+h-1}) = 1$ или $v(x_{j+h}) = 1$, что доказывает утверждение.
 
-Расчёт характеристики интервала с невычислимой и вычислимой точками производить по правилам работы с граничным интервалом, т.е.
-\begin{equation}\label{eq15} 
-R(i)=2\Delta _i-4 \frac {(z_i-z^*)}{r \mu},\text{ if } v(x_{i-1})=-1  \text{ or } v(x_{i-1})=0, v(x_i)=1,
-\end{equation}
-\begin{equation}\label{eq16} 
-R(i)=2\Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu},\text{ if } v(x_{i-1})=1, v(x_i)=-1 \text{ or } v(x_i)=0.
-\end{equation}
+Установим условия сходимости для АГП-Н и покажем, что расширение алгоритма вычисления характеристики не рождает предельных точек другого типа по сравнению с предельными точками классического АГП, кроме
+%предельных точек областей невычислимости.
 
-\begin{figure}
-\includegraphics[width=\textwidth]{pic/fig2.png}
-\caption{Примеры распределения поисковых испытаний в областях невычислимости при втором подходе к импутации значений} \label{fig2}
-\end{figure}
+\textbf{Теорема (о сходимости АГП-Н)}.
+Пусть $\overline{y}=y(\overline{x})$ есть предельная точка последовательности испытаний $\{y^k\}=\{y(x^k)\}, \{x^k\} \subset [0,1]$, порождаемой АГП-Н при минимизации функции $\phi(y), y \in D$, для которой выполнены следующие условия:
+\begin{enumerate}
+\item{$\phi(y)$ определена и вычислима в области $Q = D \backslash I$ положительного объёма, где $I$ -- область невычислимости;}
+\item{$\phi(y)$ допускает липшицево с константой $L$ продолжение $\overline{\phi}(y)$ на область $D$, т.е.
+\[
+\phi(y(x))=\overline{\phi}(y(x)), y(x) \in Q.
+\]
+}
+\end{enumerate}
+Тогда если на некотором шаге поиска выполнено условие
+\[
+r\mu > 2^{3-1/N}L\sqrt{N+3},
+\]
+то любая точка глобального минимума $y^* \in Q$ является предельной точкой последовательности $\{y^k\}$ и любая точка $\overline{y}$ является либо точкой глобального минимума функции $\phi(y), y \in Q$, либо принадлежит области невычислимости $\overline{y} \in I$.
 
-Данный подход частично решает выявленные проблемы первого подхода, однако при усложнении тестовых задач не решает проблему избыточных испытаний и требует значительного увеличения параметра надёжности $r$.
-Примеры для одномерных задач с использованием первых двух подходов представлены на рис.~\ref{fig1} и рис.~\ref{fig2}, где черным показаны испытания, попавшие в область с неопределенными значениями, а зелёным вычислимые точки области поиска.
+\textbf{Доказательство}.
+Во-первых, докажем, что для любой предельной точки $\overline{y}=y(\overline{x}), \overline{x} \in [0,1]$ существует бесконечная последовательность вложенных интервалов, соответсвующих упорядоченным по возрастанию номерам итераций $k_i$, $\{(x_{t-1}, x_t)\}, t = t(k_i), i = 1, 2, ...$, удовлетворяющая условиям:
+\[
+\lim_{t \to \infty}{\Delta_t} = 0,
+\]
+\[
+\overline{x}=\bigcap_{i = 1}^{\infty}{(x_{t-1}, x_t)},
+\]
+\[
+R(t(k_i))>0, i = 1, 2, ... ,
+\]
+\[
+\lim_{i \to \infty}{R(t(k_i))} = 0,
+\]
+где $\Delta_t = (x_t - x_{t-1})^{1/N}$, $R(t)$ и $t$ -- из Правила 5 раздела 3.1.
 
-\textbf{Третий подход.} Восстанавливать значения целевой функции в точках для интервалов с двумя невычислимыми точками как среднее между соседними точками, т.е.
-и использовать соответствующее вычислительное правило из формулы (\ref{eq9}), считая точки вычислимыми.
-Если соседние точки также являются невычислимыми, использовать формулу (\ref{eq14}).
-В случае интервала с одной невычислимой точкой производить по формулам (\ref{eq15}) и (\ref{eq16}).
+Так как $\overline{x}$ -- предельная точка, то существует последовательность номеров итераций ${k_i}, i = 1, 2, ...
+$, для которой $x^{k_i+1} \in (x_{t-1}, x_{t})$ и $\overline{x} \in (x_{t-1}, x_{t})$, $t = t(k_i)$, т.е.
+в интервал, содержащий предельную точку $\overline{x}$, попадает точка очередного испытания $x^{k_{i}+1}$.
+Из правил вычисления новой точки (\ref{eq10})--(\ref{eq11}), а также оценки (\ref{eq7}) получаем
+оценку для максимальной длины очередного поискового интервала
+%\[
+%x_t-x^{k_i+1} \leq \frac{x_t-x_{t-1}}{2} + \text{sign}(z_t-z_{t-1})\frac{1}{2r}\Delta_t^N \leq \frac{r + 1}{2r}(x_t-x_{t-1}),
+%\]
+%аналогично для $x^{k_i+1} - x_{t-1}$, тогда справедлива оценка
+\[
+\max\{x_t-x^{k_i+1},x^{k_i+1} - x_{t-1}\} \leq \gamma(x_t - x_{t - 1}),
+\]
+где коэффициент $\gamma = \frac{r+1}{2r} < 1$. Отсюда следует, что $\Delta_t \to 0$ при $i \to \infty$ и $\overline{x}=\bigcap_{i = 1}^{\infty}{(x_{t-1}, x_t)}$.
 
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{pic/fig3.png}
-\caption{Импутация невычислимых значений средним соседних точек. Подход \No~3} \label{fig3}
-\end{figure}
+%В предположении, что на каждом шаге $k \geq 1$ существует некоторый интервал $(x_{j-1}, 
+Предположим, что на каждом шаге $k \geq h$ существует некоторый интервал $(x_{j-1}, x_{j})$, в одной из точек которого достигается текущее минимальное значение $z^*$, т.е.
+$z^* = z_j$ или $z^* = z_{j-1}$.
+Тогда справедливо соотношение $z_j + z_{j-1}-2z^* = |z_j-z_{j-1}| = \mu_j \Delta_j \leq \mu \Delta_j$ и, если обе граничные точки вычислимы $v(x_{j-1}) = v(x_j) = 1$, то из (\ref{eq9}) следует
+\[
+R(j) = \Delta_j + \frac{(\mu_j \Delta_j)^2}{(r\mu)^2 \Delta_j}-2\frac{\mu_j \Delta_j}{r\mu} = (\Delta_j-\frac{\mu_j\Delta_j}{r\mu})^2 \geq (1-1/r)^2 \Delta_j > 0,
+\]
+а если $v(x_{j-1}) \neq v(x_j)$, то $R(j)=2\Delta_j>0$.
+Таким образом, на каждом шаге поиска $k \geq h$ существует интервал со сторого положительной характеристикой.
+Значит и характеристики лучших интервалов с номерами $t = t(k_i)$ будут положительными, т.е.
+$R(t(k_i))>0, i = 1, 2, ...
+$.
 
-Существенными минусами будет не использование информации о координатах, а также редкость работы добавленного вычислительного правила за счёт поиска значений на замену только среди непосредственных соседей.
+Из правил вычисления характеристик интервалов (\ref{eq9}) следует, что если $v(x_{t-1}) = v(x_t) = 1$, то для характеристики лучшего интервала справедлива оценка сверху
+\[
+R(t) \leq \Delta_t + \frac{(\mu \Delta_t)^2}{(r\mu)^2 \Delta_t} = (1+1/r^2) \Delta_t.
+\]
+В случае если $v(x_{t-1}) = v(x_t) = 0$ или $v(x_{t-1}) = 0$ и $v(x_t) = -1$ или $v(x_{t-1}) = -1$ и $v(x_t) = 0$, имеет место строгое равенство $R(t) = (1-1/r)^2\Delta_t$, в остальных случаях $R(t) \leq 2\Delta_t$.
+Из полученных оценок и в силу уже доказанного ранее свойства $\lim_{i \to \infty}{\Delta_t} = 0$ следует, что $\lim_{i \to \infty}{R(t(k_i))}=0, i = 1, 2, ...$.
 
-\textbf{Четвёртый подход.} Восстанавливать значения целевой функции в точках для интервалов с двумя невычислимыми точками как среднее между ближайшими вычислимыми точками с учётом положения координат, т.е.
-\begin{equation}\label{eq17} 
-%\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
-\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \rho(x,x')}{\rho(x',x'') },
-\end{equation}
-где $z'$ и $z''$ -- значения целевой функции в ближайших к интервалу $(x_{i-1},x_i)$ вычислимых точек $x'$ и $x''$, а $\rho(x_1,x_2) =  |x_1 - x_2|^{1/N}$.
-В остальном подход идентичен третьему.
-Формула (\ref{eq14}) в данном случае будет применяться лишь тогда, когда невычислимая область находится по соседству с граничной точкой.
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{pic/fig4.png}
-\caption{Импутация невычислимых значений средним соседних точек. Подход \No~4} \label{fig4}
-\end{figure}
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{pic/fig5.png}
-\caption{Импутация невычислимых значений средним между минорантой и мажорантой. Подход \No~5} \label{fig5}
-\end{figure}
+Во-вторых, покажем, что если $y^* = y(x^*) \in Q$ -- точка глобального минимума, то $y^*$ является предельной точкой последовательности $\{y^k\}$.
 
-\textbf{Пятый подход.} Чтобы максимально использовать поисковую информацию и предположения о виде функции, можно восстанавливать значения целевой функции в точках с учётом условия Гёльдера.
+Пусть $x^* \in (x_{j-1}, x_{j})$, тогда она может быть представлена в виде
+\[
+x^* = \beta x_{j-1}+(1-\beta)x_j, 0 \leq \beta \leq 1,
+\]
+а в силу условия Гелдера (\ref{eq4}) справедливы оценки
+\[
+z_j=\phi(y(x_j)) \leq \phi(y(x^*)) + 2L\sqrt{N+3}(x_j - x^*)^{1/N}, \text{ если } v(x_j) = 1,
+\]
+\[
+z_{j-1}=\phi(y(x_{j-1})) \leq \phi(y(x^*)) + 2L\sqrt{N+3}(x^* - x_{j-1})^{1/N},\text{ если } v(x_{j - 1}) = 1.
+\]
+Тогда для интервала с двумя вычислимыми граничными точками имеет место соотношение
+\[
+z_j+z_{j-1} \leq 2\phi(y(x^*)) + 4L\sqrt{N+3}((x^* - x_{j-1})^{1/N}+(x_j-x^*)^{1/N}) =
+\]
+\[
+= 2\phi(y(x^*)) + 4L\sqrt{N+3}(((1-\beta)(x_j-x_{j-1}))^{1/N}+(\beta (x_j-x_{j-1}))^{1/N}) \leq
+\]
+\[
+\leq 2\phi(y(x^*)) + 4\Delta_j L\sqrt{N+3}\max_{0< \beta < 1}{\{(1-\beta)^{1/N}+\beta^{1/N}\}} = 2\phi(y(x^*)) + 2^{2-\frac{1}{N}}\Delta_j L\sqrt{N+3}.
+\]
 
-\textit{Шаг 1.} Найти крайние точки пересечения миноранты и мажоранты
-\begin{equation}\label{eq18} 
-\hat{x}',\hat{x}''=\frac {x'+x''}{2}\pm \text{sign}(z''-z')\cdot \frac {1}{2r} \cdot {\left(\frac {z''-z'}{\mu}\right)}^N,
-\end{equation}
+Из выведенных соотношений и правил (\ref{eq9}) следует, что характеристики интервалов будут строго положительными:
+\[
+R(j) =(1-\frac{1}{r})^2\Delta_i>0, \text{ если } v(x_j) = v(x_{j-1}) = 0,
+\]
+\[
+R(j) \geq \Delta_j-2\frac{2\phi(y(x^*)) + 2^{2-\frac{1}{N}}\Delta_j L\sqrt{N+3}-2z^*}{r\mu} \geq \Delta_j(1-\frac{2^{3-1/N}L\sqrt{N+3}}{r\mu}) + 4\frac{(z^*-\phi(y(x^*)))}{r\mu} \leq
+\]
+\[
+\leq \Delta_j(1-\frac{2^{3-1/N}L\sqrt{N+3}}{r\mu})>0, \text{ если } v(x_j) = v(x_{j-1}),
+\]
+\[
+R(j) \geq 2\Delta_j-4\frac{\phi(y(x^*))+2\Delta_jL\sqrt{N+3}-z^*}{r\mu} \geq 2\Delta_j(1-\frac{4L\sqrt{N+3}}{r\mu}) + 4\frac{4(z^*-\phi(y(x^*)))}{r\mu}\geq
+\]
+\[
+\geq 2\Delta_j(1-\frac{4L\sqrt{N+3}}{r\mu})>0,\text{ если } v(x_j) \neq v(x_{j-1}),
+\]
+если будет выполнено условие $r\mu > 2^{3-1/N}L\sqrt{N+3}$.
 
-\textit{Шаг 2.} Восстановить значения целевой функции по правилу
-\begin{equation}\label{eq19} 
-\tilde{z}(x)=
-  \begin{cases}
-    z', & {\quad x \leq \hat{x}',}\\
-    %\frac {z'+z''}{2}- \frac {r \mu}{2} (\| x - x' \| + \| x - x'' \|), & {\quad \hat{x}' < x < \hat{x}'',}\\
-    \frac {z'+z''}{2}- \frac {r \mu}{2} (\rho(x,x')  + \rho(x,x'')), & {\quad \hat{x}' < x < \hat{x}'',}\\
-    z'',  & {\quad x \geq \hat{x}'',}
-  \end{cases}
-\end{equation}
-где $\rho(x_1,x_2) =  |x_1 - x_2|^{1/N}$.
-Значение $\tilde{z}(x)$ может быть интерпретировано как среднее значение между минорантой и мажорантой для функции, удовлетворяющей условию Гёльдера.
+Значит в интервал $(x_{j-1}, x_{j})$ попадёт точка одного из следующих испытаний.
+Повторяя данные рассуждения, приходим к выводу, что правила алгоритма порождают вложенную последовательность сжимающихся к $x^*$ интервалов, т.е.
+$x^*$ -- предельная точка и $\lim_{k \to \infty}{\{x^k\}} = x^*$.
 
-Приведём описание модификации алгоритма глобального поиска на случай не всюду вычислимой функции на основе данного подхода.
+В-третьих, покажем, что если $\overline{y} = y(\overline{x})$ -- предельная точка последовательности $\{y^k\}$, то $\overline{y}$ является либо точкой глобального минимума функции $\phi(y), y \in Q$ , либо принадлежит области невычислимости $\overline{y} \in I$.
 
-\subsection{Описание модифицированного алгоритма глобального поиска}
-Алгоритм остаётся прежним с внесением в него следующих небольших дополнений для вычислительных формул на случай попадания в невычислимые области поиска:
-
-\begin{enumerate}[itemsep=0pt,parsep=2pt,topsep=2pt,partopsep=0pt]
-  \item Индекс $v_i=v(x_i)$ определяется по формуле (\ref{eq12}).
-  \item При вычислении текущей нижней оценки для константы Гельдера $K$ в формуле (\ref{eq7}) использовать только интервалы с двумя вычислимыми точками.
-  \item Текущее лучшее значение целевой функции $z_v^*$ определяется для разных значений индекса
-\begin{equation}\label{eq20} 
-z_v^*=
-  \begin{cases}
-    -\varepsilon _r, & {\quad v=0 ,}\\
-    \min \left\{ f(x_i): v(x_i)=1, 0 < i < k \right\}, & {\quad v=1 ,}
-  \end{cases}
-\end{equation}
-где $\varepsilon _r > 0$ -- параметр метода.
-
-  \item Для каждого интервала $(x_{i-1}, x_i), 1 \leq i \leq k$, the вычисление характеристики $R(i)$ происходит согласно следующим вычислительным формулам.
-
-        \textit{Случай 1.} Для интервалов с \textbf{двумя вычислимыми точками} или \textbf{вычислимой и граничной точками} применить классические правила алгоритма глобального поиска
-\begin{equation}\label{eq21} 
-R(i)=
-  \begin{cases}
-     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z_1^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1,}\\
-    2 \Delta _i-4 \frac {(z_i-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1,}\\
-    2 \Delta _i-4 \frac {(z_{i-1}-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
-  \end{cases}
-\end{equation}
-        \textit{Случай 2.} Для интервалов с \textbf{граничной и невычислимой точками} произвести импутацию значения для невычислимой точки на основе соседней вычислимой точки и применить формулу
-\begin{equation}\label{eq22} 
-R(i)=
-  \begin{cases}
-    2 \Delta _i-4 \frac {(z_{i+1}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=0,}\\
-    2 \Delta _i-4 \frac {(z_{i-2}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=-1.}
-  \end{cases}
-\end{equation}
-        \textit{Случай 3.} Для интервалов с \textbf{невычислимой и вычислимой точками} произвести импутацию значения для невычислимой точки на основе соседней вычислимой точки и применить формулу
-\begin{equation}\label{eq23} 
-R(i)=
-  \begin{cases}
-     \Delta _i+\frac {{(z_i-z_{i-2})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-2}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=1,}\\
-     \Delta _i+\frac {{(z_{i+1}-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_{i+1}+z_{i-1}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0.}
-  \end{cases}
-\end{equation}
-        \textit{Случай 4.} Для интервалов с \textbf{двумя невычислимыми точками} произвести импутацию значений для невычислимой точки на основе правила (\ref{eq19}) и применить формулу
-\begin{equation}\label{eq24} 
-R(i)=\Delta _i+\frac {{(\tilde{z}_i-\tilde{z}_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {\tilde{z}_i+\tilde{z}_{i-1}-2z_1^*}{r \mu}, \; v(x_{i-1})=v(x_i)=0.
-\end{equation}
-        \textit{Случай 5.} В случае невозможности проведения любой импутации использовать формулу
-\begin{equation}\label{eq25} 
-R(i)=\Delta _i \cdot {\left( 1-\frac {1}{r} \right)}^N+z_0^*.
-\end{equation}
-  \item Если максимальную характеристику имеют несколько интервалов, отдавать предпочтение вычислимым интервалам с минимальным номером.
-  \item Если лучшим интервалом оказался интервал с двумя невычислимыми точками, очередное испытание проводить в серединной точке интервала.
-\end{enumerate}
-
-\begin{nmAlgorithm}{pril:prog:alg1}
-      {Функция $g$ (вычисление координат точки $v$ по ее порядковому номеру $k$)}
-      {Function $g$ (calculating the coordinates of the point $v$ by its ordinal $k$)}
-     \algLine{\algFUNC $g(k,d,p)$}
-     \algLine[1]{$u_{n-1} \ceq \left\lfloor k/\left(d-1\right)^{n-2}\right\rfloor$}
-     \algLine[1]{$u_n \ceq u_{n-1}$}
-     \algLine[1]{$k\ceq k\mod (d-1)^{n-2}$}
-     \algLine[1]{\algFOR $j=\left(n-3\right)\ldots0$ \algDO}
-     \algLine[2]{$u_j\ceq \left\lfloor k/(d-1)^j\right\rfloor+1$}
-     \algLine[2]{$k\ceq k\mod(d-1)^j$}
-     \algLine[1]{\algENDFOR}
-     \algLine[1]{\textbf{return} $(v_1,\dots,v_n)$}
-     \algLine{\algENDFUNC}
-\end{nmAlgorithm}
-
-\subsection{Условия сходимости}
-
-Установим условия сходимости для модификации алгоритма и покажем, что расширение алгоитма вычисления характеристики не рождает предельных точек другого типа по сравнению с предельными точками классического АГП.
-
-\textbf{Теорема}. Формулировка теоремы.
-
-\textbf{Доказательство}. Доказательство теоремы.
-
-\section{Результаты вычислительных экспериментов}
-
-Вычислительные эксперименты проводились на машине с процессором Intel\textregistered\ Core\texttrademark\ i7-10750H CPU @ 2.60GHz с использованием разработанного программного комплекса, реализующего описанную модификацию из Раздела 3.
-Для генерации серии задач использовался GKLS-генератор, описанный в [\ref{rfa:rulit:Gaviano2003}].
-Он позволяюет порождать задачи многоэкстремальной оптимизации с заранее известными свойствами.
-Данные задачи были дополнены областями невычислимости разных видов.
-
-Каждая серия экспериментов состоит из 100 задач GKLS со всюду вычислимой целевой функцией (original), 100 задач с невычислимостью на границе области поиска (round) и 100 задач со случайно заданными областями невычислимости (random).
-
-\begin{figure}[h]
-\includegraphics[width=\textwidth]{pic/fig6.png}
-\caption{Примеры тестовых задач с \No~19, 25, и 35 из первой серии экспериментов} \label{fig6}
-\end{figure}
-
-В первой серии экспериментов проверялась корректность работы модификации и выявлялись особенности проведения испытаний в областях невычислимости.
-В эксперименте использовались задачи размерности $N=2$, параметр надежности был задан $r=5.5$, а точность $\varepsilon=0.01$.
-Результаты экспериментов представлены в табл.~\ref{tab1}.
-
-На рис.~\ref{fig6} приведены примеры работы модификации с несколькими задачами серии.
-На рисунках синими точками отмечены испытания в вычислимых областях, красным найденный глобальный минимум задачи, а черным -- точки испытаний, проводившихся в области невычислимости.
-Также для наглядности области невычислимости окрашены в серый цвет.
-
-Так как модификация с классической серией задач GKLS работает согласно базовому алгоритму глобального поиска, можно провести ее сравнение с классическим АГП.
-Так, для модификации сохранено общее поведение АГП и наблюдаются те же области сгущения испытаний вблизи локальных минимумов, что объясняется срабатыванием в большинстве случаев именно классических вычислительных правил (\ref{eq21}), в табл.~\ref{tab1} видно, что число испытаний, проводимых в невычислимых областях не велико.
-Общее число испытаний для классических задач и задач с невычислимостями соизмеримо, метод не проводит избыточных итераций в неперспективных областях благодаря расчётным формулам, использующим информацию о ближайших вычислимых точках.
-
-\begin{table}[h]
-\caption{Сравнение модификации алгоритма с классической версией алгоритма глобального поиска}\label{tab1}
-\centering
-\begin{tabular}{|l|c|c|c|}
-\hline
- &  { original }  & { round }  & { random }  \\
-\hline
-Среднее число испытаний & 781.7	& 854.9 & 839.0 \\
-\begin{tabular}{@{}l@{}} {Среднее число испытаний} \\  { в областях невычислимости } \end{tabular}  & - & 185.8 & 42.3 \\
-{Число решенных задач} & 100 / 100 & 100 / 100  &  100 / 100 \\
-\hline
-\end{tabular}
-\end{table}
-
-Использование правила (\ref{eq25}) формирует сетку испытаний внутри граничных областей невычислимости, что хорошо заметно на задаче \No~35.
-При этом у границы невычислимой области близкой к потенциальному минимума формируется сгущение точек испытаний при попытке подойти к нему с разных сторон благодаря вычислительным правилам (\ref{eq22})--(\ref{eq24}), что наблюдается во всех приведенных в качестве примеров задачах.
-
-Решаемость задач 100\%, однако для серии с граничной областью потребо-валось дорешивание одной задачи с немного большим значением $r=6.0$, для серии со случайными областями то же значение $r$ использовалось для решения двух задач.
-
-\begin{table}
-\caption{Серия задач разной размерности}\label{tab2}
-\centering
-\begin{tabular}{|l|l|c|c|c|}
-\hline
- \begin{tabular}{@{}c@{}} {N} \\ {}\end{tabular} &
- \begin{tabular}{@{}c@{}} {Тип} \\ {тестовых задач} \end{tabular} &
- \begin{tabular}{@{}c@{}} {Среднее число испытаний} \\ {} \end{tabular} &
- \begin{tabular}{@{}c@{}} {Среднее число испытаний} \\ {в области невычислимости} \end{tabular} &
- \begin{tabular}{@{}c@{}} {Число решенных задач} \\ {}\end{tabular} \\
-\hline
-$3$ & original & 12251.1 & - & 100 / 100 \\
-$ $ & round & 12201.8 & 4763.6 & 100 / 100 \\
-$ $ & random & 12210.7 & 63.4 & 100 / 100 \\
-\hline
-$4$ & original & 25144.5 & - & 100 / 100 \\
-$ $ & round & 29840.3 & 15004.4  & 100 / 100 \\
-$ $ & random & 25287.5 & 26.4 & 100 / 100 \\
-\hline
-$5$ & original & 56725.9 & - & 100 / 100 \\
-$ $ & round & 47840.1 & 24109.2 & 100 / 100\\
-$ $ & random & 57874.8 & 82.6 & 100 / 100\\
-\hline
-\end{tabular}
-\end{table}
-
-Во второй серии экспериментов проводилось решение серий задач разных размерностей.
-В эксперименте использовались задачи размерности $N=3,4,5$, параметр надежности был задан $r=4.5$, а точность $\varepsilon=0.02$.
-Глобальный минимум считался найденным, если алгоритм генерировал точку испытания в $\varepsilon$-окрестности глобального минимума.
-Результаты выполненных экспериментов представлены в табл.~\ref{tab2}.
-
-Как показывают результаты выполненных экспериментов, метод отлично справляется со случайными попаданиями в области невычислимости, разбросанными по области поиска случайно, а в случае больших граничных областей пытается провести полноценное исследование, не проводя при этом слишком много испытаний.
+Пусть $\overline{y} = y(\overline{x})$ -- предельная вычислимая точка, $\overline{y} \in Q$ и она не является точкой глобального минимума.
+%, т.е. $\phi(y(\overline{x})) > \phi(y(x^*))$.
+Тогда существует последовательность интервалов с номерами $t = t(k_i), i = 1, 2, ...$, стягивающихся к точке $y(\overline{x})$ и имеющих положительные характеристики, и имеет место
+\[
+\phi(y(\overline{x})) \geq z^* + \delta > \phi(y(x^*)), \delta > 0.
+\]
+Тогда для очередного лучшего интервала при достаточно больших t справедливо $z_{t-1} \geq z^* + \delta$, если $v(x_{t-1}) = 1$, и $z_{t} \geq z^* + \delta$, если $v(x_{t}) = 1$. 
+Из правил вычисления характеристик следует
+\[
+R(t) < \Delta_t-\frac{4\delta}{r\mu} < 0, \text{ если } v(x_{t}) = 1 \text{ и } v(x_{t-1}) = 1,
+\]
+\[
+R(t) < 2\Delta_t-\frac{4\delta}{r\mu} < 0, \text{ если } v(x_{t}) \neq v(x_{t-1}),
+\]
+так как, как было показано ранее $\lim_{t \to \infty}{\Delta_t} = 0$, что противоречит с доказанным ранее свойством положительности харакетристик стягивающихся интервалов $R(t) > 0$.
+Значит наше предположение неверно и либо предельная точка $\overline{y} = y(\overline{x})$ является точкой глобального минимума, либо функция в ней является невычислимой и алгоритмом порождается последовательность невычислимых стягивающихся интервалов с положительной характеристикой, вычисляемой по правилу $R(t) =(1-\frac{1}{r})^2\Delta_t$.
 
@@
 В таких подобластях невозможно корректно провести численное моделирование и вычислить значение целевой функции.
-Появление данной постановки обусловлено частой проблемой численной нестабильности исследуемых моделей прикладных задач.
 
@@
 
-Исследуемая в данной работе проблема является частичным аналогом уже давно существующей в Data Science проблемы пропущенных значений.
-В качестве решения специалистами в данной сфере были предложены различные подходы к импутации (восстановлению) пропущенных значений.
-%Эти идеи нашли своё отражение при разработке нового метода.
-
-Было проведено исследование применимости уже существующих подходов, дополненных различными вариантами учета липшицевости функции и использующих информационно-статистическую теорию.
-В результате удалось разработать универсальный эффективный алгоритм для данного класса задач.
-Построенный алгоритм является развитием одного из эффективных детерминированных методов решения задач липшицевой глобальной оптимизации – информационно-статистического алгоритма глобального поиска.
-В рамках модификации была произведена адаптация вычислительных правил алгоритма с внедрением алгоритма импутации значений.
-
-Были проведены численные эксперименты на сериях многоэкстремальных задач GKLS размерностей $N = 2,3,4,5$.
-Результаты продемонстрировали надежность и эффективность предложенного подхода.
-Модификация отлично справляется со случайными попаданиями в подобласти невычислимости, разбросанными по области поиска.
-В случае больших граничных областей невычислимости метод всё также эффективно решает задачу, проводя полноценное исследование, в том числе, в невычислимых областях.
+В статье было приведено описание одной из исследованных ранее модификаций алгоритма глобального поиска для решения такого класса задач.
+Приведены теоретические обоснования сходимости алгоритма.
+Теоретические результаты полностью соответсвуют картине, полученой в ходе экспериментальных исследований эффективности алгоритма.
+Разработанная модификация обладает недостатком в виде избыточных испытаний в области невычислимости (в силу порождения алгоритмом предельных точек в данных областях).
+Данного явления можно избежать с использованием более сложной модификации, восстанавливающей значения в невычислимых точках, основываясь на предположении о липшивости целевой функции (что было отражено в экспериментальных исследованиях).
+Теоретическое обоснование данной модификации является объектом будущих исследований.
 
@@
 \doi{10.1007/s00500-020-05078-1}.
+}
+
+\rlitem{rfa:rulit:Usova2024}{%
+\textit{M.~A.~Usova, K.~A.~Barkalov} 
+An Algorithm for Finding the Global Extremum of a Partially Defined Function~// Communications in Computer and Information Science. 2024. {\bf 1914}. 147--161.
+\doi{10.1007/978-3-031-52470-7_13}.
 }
@@
 }
+
+\elitem{rfa:enlit:Usova2024}{%
+M.~A.~Usova, K.~A.~Barkalov, ``An Algorithm for Finding the Global Extremum of a Partially Defined Function'', Communications in Computer and Information Science. {\bf 1914}, 147--161 (2024).
+\doi{10.1007/978-3-031-52470-7_13}.
+}
+
 %2

### 2025-02-03 2025/Problems of Informatics/article.tex
added text of paper to Problems of Informatics

\documentclass[a4paper,12pt,russian]{article}
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[russian]{babel}
\usepackage[a4paper,left=2cm, right=2cm, top=2cm, bottom=2cm]{geometry}
\usepackage{graphicx}
\graphicspath{{figures/}}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{caption} 
\linespread{1.3}

% дополнительно
%\usepackage{algorithm}
%\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{listings}

\begin{document}

\begin{center}
	\textbf{AN ALGORITHM FOR SOLVING GLOBAL OPTIMIZATION\break
    PROBLEMS WITH PARTIALLY DINEFED FUNCTIONS}
\end{center}

\begin{center}
{M.A.~Usova, I.G.~Lebedev, A.A.~Shtanyuk, K.A.~Barkalov}
\end{center}

\begin{center}
{Lobachevsky State University, 603022, Nizhny Novgorod, Russia}
\end{center}
\begin{small}
% аннотация англ.
The paper discusses the problem of finding a global minimum of a function that may be multiextremal, non-differentiable, given in the form of a ``black box'' and computable only in some part of the search domain. In this case, each computation of the objective function value at some point of the feasible domain may require significant computing resources. The objective function is assumed to satisfy the Lipschitz condition.
The existence of subdomains where the objective function is undefined can be interpreted as the existence of some hidden, a priori unknown constraints of the problem. Such an effect can occur when solving applied problems due to the specific features of the optimized object or modeling method (for example, numerical instability of the method for some combination of parameters). In such subdomains, it is impossible to correctly perform numerical modeling and estimate the value of the objective function. A special case is the problem of tuning hyperparameters of artificial intelligence (AI) and machine learning (ML) methods. In these problems, the efficiency (in a given metric) of the resulting solution for different values of hyperparameters can be quite different. Moreover, the computational complexity of tuning makes the exhaustive search practically inapplicable. These features required the development of various methods for intelligent automatic tuning of hyperparameters. At the same time, the problem of existence of unacceptable combinations of hyperparameters, which was not particularly important in the case of manual tuning, has become a stumbling block for many frameworks using intelligent tuning methods.
Our goal is to develop an algorithm in which the absence of a value of the objective function at some point should not lead to its incorrect operation. The proposed approach to solving such problems extends the information-statistical global search algorithm (GSA). This algorithm solves multidimensional problems by reducing them to one-dimensional optimization problems using space-filling curves (Peano curves).
In this study, new rules for processing search intervals with no objective function values at their boundary points were proposed. The calculation of the characteristic of an interval with two non-computable points is based on the interval length. To minimize the number of redundant trials in subdomains where the function is not defined, we used a special method parameter $\alpha$ that regulates the number of trial points in these subdomains. The calculation of the characteristic of an interval with non-computable and computable points is performed according to the rules for working with a boundary interval. The article provides a detailed description of a modified global search algorithm for solving the class of problems under consideration.
The implementation of the global search algorithm for the case of a not-everywhere computable objective function (GSA-N) was based on the iOpt open source framework of intelligent optimization methods.
To carry out the experiments, a generator of test problems with hidden constraints GKLS-HC was developed. It is based on the GKLS generator, which allows generating multi-extremal functions with specified properties (number of minima, their regions of attraction, etc.). In the GKLS-HC generator, these functions were spoiled by areas of non-computability in the form of ellipsoids (the coordinates of centers and radii were generated randomly).
The experimental results presented in the paper, obtained on a series of GKLS-HC test problems, demonstrate the reliability of global search and the efficiency of the developed algorithm. By adjusting the parameter $\alpha$, it was possible to achieve the same speed of GSA-N operation as the basic GSA.
The paper also considered the behavior of global optimization algorithms of the \textit{scipy.optimize} library when solving problems with non-computable domains. In this study, the differential evolution and brute force methods showed the worst results, failing to solve this type of problem at all. The DIRECT and SHGO methods, although they solved the problems, were less effective than the developed GSA-N.
In addition, the paper presents the results of numerical experiments obtained when tuning the hyperparameters of the LinearSVC method used to solve the classification problem on the well-known Iris dataset.

%Here is the text of the extended annotation in English, ranging from 4000 to 8000 characters.
%В статье обсуждается проблема поиска глобального минимума функции, которая может быть многоэкстремальной, недифференцируемой, более того, заданной в форме «черного ящика» и вычислимой лишь в некоторой части области поиска. При этом каждое вычисление функции в некоторой точке допустимой области может требовать значительных вычислительных ресурсов. О целевой функции известно лишь что она удовлетворяет условию Липшица.
%Наличие подобластей, в которых целевая функция является неопределенной, можно интерпретировать как наличие некоторых скрытых, заранее неизвестных ограничений задачи. Такой эффект может возникать при решении прикладных задач в силу особенностей оптимизируемого объекта или метода моделирования (например, численная нестабильность метода при определенном сочетании параметров). В таких подобластях невозможно корректно провести численное моделирование и оценить значение целевой функции. Частным случаем является задача настройки гиперпараметров методов искусственного интеллекта (ИИ) и машинного обучения (МО). Во многих случаях разница в эффективности (в заданной метрике) получаемого решения при разных значениях гиперпараметров может быть достаточно значительной. При этом вычислительная сложность настройки делает невозможным использование полного перебора. Данные особенности определили необходимость разработки различных методов интеллектуального автоматического подбора гиперпараметров. При этом проблема возникновения недопустимых комбинаций гиперпараметров, которая не имела особого значения в случае «ручной» настройки, для многих фреймворков, использующих методы интеллектуальной настройки, стала камнем преткновения.
%Нашей целью является разработка алгоритма, в котором отсутствие значения целевой функции в некоторой точке не должно приводить к его некорректной работе. Предлагаемый подход к решению такого рода задач является расширением информационно-статистического алгоритма глобального поиска (АГП). Данный алгоритм предполагает решение многомерных задач посредством их сведения к задачам одномерной оптимизации при помощи кривых, заполняющих пространство (кривых Пеано). В рамках данного исследования были предложены новые правила обработки поисковых интервалов, в граничных точках которых отсутствуют значения целевой функции. Расчёт характеристики интервала с двумя невычислимыми точками производится, основываясь на длине интервала. Для минимизации количества избыточных испытаний в областях, в которых функция не определена, при таком подходе используется специальный параметр метода, позволяющий регулировать число точек испытаний в области невычислимости.  Расчёт характеристики интервала с невычислимой и вычислимой точками производится по правилам работы с граничным интервалом. В статье изложено подробное описание модифицированного алгоритма глобального поиска для решения исследуемого класса задач.
%Программная реализация алгоритма глобального поиска на случай невсюду вычислимой целевой функции (АГП-Н) была выполнена на базе фреймворка методов интеллектуальной эвристической оптимизации iOpt открытым исходным кодом.
%Для проведения экспериментов был разработан генератор задач со скрытыми ограничениями GKLSHiddenConstraint, взявший за основу генератор GKLS, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами. Данные функции были дополнены сгенерированными областями невычислимости в форме эллипсоидов (координата центры и радиусы генерировались случайно).
%Приведенные в данной статье результаты экспериментов на серии тестовых задач GKLSHiddenConstraint демонстрируют надёжность глобального поиска и эффективность разработанного алгоритма. За счёт регулировки параметра $\alpha$ получилось добиться такой же скорости сходимости АГП-Н, как у базового АГП. 
%Также в статье было изучено поведение алгоритмов глобальной оптимизации библиотеки \textit{scipy.optimize} при решении задач с невычислимыми областями. В рамках данного исследования методы differential\_evolution и brute показали наихудшие результаты, не справившись с решением такого вида задач совсем. Методы direct и shgo хоть и продемонстрировали хорошие результаты, оказались менее эффективны чем разработанный АГП-Н. 
%Помимо этого в статье продемонстрированы результаты численных экспериментов с прикладной задачей настройки гиперпараметров метода LinearSVC при решении задачи классификации на известном наборе данных Iris. <РЕЗУЛЬТАТЫ>.

\textbf{Keywords:} Lipschitz global optimization, multi-extremal functions, black-box functions, partially defined functions, hidden constraints, hyperparameter tuning.
\newline
\end{small}
\newline
{\large \textbf{References}}
% если есть русскоязычная нужна транслитерация неанглоязычных элементов списка литературы
\begin{enumerate}
%7
%\item \label{rfa:enlit:Grishagin2016_2}
%V.~A.~Grishagin, R.~A.~Israfilov, ``Global search acceleration in the nested optimization scheme'', AIP Conference Proceedings. {\bf 1738}, 400010 (2016). DOI: 10.1063/1.4952198.
\item \label{rfa:enlit:Jones2021}
Jones~D., Martins~J. The direct algorithm: 25 years later~// J. Glob. Optim. 2021. Vol. 79, No,~3. P. 521--566. DOI: 10.1007/s10898-020-00952-6.

%12
\item \label{rfa:enlit:PaulaviciusZilinskas2014}
Paulavi{\v c}ius~R. and {\v Z}ilinskas~J. Simplicial Global Optimization. New York: Springer, 2014. DOI: 10.1007/978-1-4614-9093-7.

%13
\item \label{rfa:enlit:Birect2020}
Paulavi{\v c}ius~R., Sergeyev~Y.D., Kvasov~D.E., {\v Z}ilinskas~J. Globally-biased {BIRECT} algorithm with local accelerators for expensive global optimization~// 
Expert Syst. Appl. 2020. Vol. 144. p. 113052. DOI: 10.1016/j.eswa.2019.113052.

%14
\item \label{rfa:enlit:Sergeyev2017}
Sergeev~Ya.D., Kvasov~D.E. Diagonalnye metody globalnoj optimizacii.  M.: Fizmatlit, 2008.

%11
\item \label{rfa:enlit:Liberti2005}
Liberti~L., Kucherenko~S. Comparison of deterministic and stochastic approaches to global optimization~// Int. Trans. Oper. Res. 2005. Vol. 12. P. 263--285.

%15
\item \label{rfa:enlit:Sergeyev2018}
Sergeyev~Y.D., Kvasov~D.E., Mukhametzhanov~M.S. On the efficiency of nature-inspired metaheuristics in expensive global optimization with limited budget~// Sci. Rep. 2018. Vol. 8, No.~1. p. 435.

%16
\item \label{rfa:enlit:Sergeyev2013}
Sergeyev~Y.D., Strongin~R.G., Lera~D. Introduction to Global Optimization Exploiting Space-Filling Curves. New York: Springer Briefs in Optimization, 2013. DOI: 10.1007/978-1-4614-8042-6.

%20
\item \label{rfa:enlit:Strongin2000}
Strongin~R.G., Sergeyev~Y.D. Global optimization with non-convex constraints. Sequential and parallel algorithms. Dordrecht: Kluwer Academic Publishers, 2000.

%10
\item \label{rfa:enlit:Kvasov2013}
Kvasov~D.E., Sergeev~Ya.D. Metody lipshicevoj globalnoj optimizacii v zadachax upravleniya~// Avtomatika i telemexanika. 2013. No. 9. S. 3--19.

%17
\item \label{rfa:enlit:Sergeyev2020}
Sergeyev~Y.D., Candelieri~A., Kvasov~D.E., Perego~R. Safe global optimization of expensive noisy black-box functions in the $\delta$-Lipschitz framework~// 
Soft Comput. 2020. Vol. 24, No.~23. P. 17715--17735. DOI: 10.1007/s00500-020-05030-3.

%4
\item \label{rfa:enlit:Dongarra2022}
Dongarra~J.J. The evolution of mathematical software~// Commun. ACM. 2022. Vol. 65, No.~12. P. 66--72. DOI: 10.1145/3554977.

%5
\item \label{rfa:enlit:Duwe2020}
Duwe~K., et al. State of the art and future trends in data reduction for high-performance computing~// Supercomput. Front. Innov. 2020. Vol. 7, No.~1. DOI: 10.14529/jsfi200101.

%19
\item \label{rfa:enlit:Stripinis2021}
Stripinis~L., Paulavi{\v c}ius~R. A new {DIRECT}-{GLh} algorithm for global optimization with hidden constraints~// Optim. Lett. 2021. Vol. 15, No.~6. P. 1865--1884.
DOI: 10.1007/s11590-021-01726-z.

%1
\item \label{rfa:enlit:Audet2022}
Audet~C., Batailly~A., Kojtych~S. Escaping unknown discontinuous regions in blackbox optimization~// SIAM J. Optim. 2022. Vol. 32, No.~3. P. 1843--1870. DOI: 10.1137/21m1420915.

%3
\item \label{rfa:enlit:Candelieri2019}
Candelieri~A. Sequential model based optimization of partially defined functions
under unknown constraints~// J. Glob. Optim. 2019. Vol. 79, No.~2. P. 281--303. DOI: 10.1007/s10898-019-00860-4.

%18
\item \label{rfa:enlit:Sergeyev2003}
Barkalov~K.A., Strongin~R.G. Metod globalnoj optimizacii s adaptivnym poryadkom proverki ogranichenij~// Zhurn. vy`chisl. matem. i matem. fiz. 2002. T.~42. No.~9. S.~1338--1350.

%21
\item \label{rfa:enlit:Strongin2020}
Strongin~R.G., Barkalov~K.A., Bevzuk~S.A. Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints~// 
Soft Comput. 2020. Vol. 24, No.~16. P. 11853--11865. DOI: 10.1007/s00500-020-05078-1.

\item \label{rfa:enlit:Usova2024}
Usova~M.A., Barkalov~K.A. An Algorithm for Finding the Global Extremum of a Partially Defined Function~// Communications in Computer and Information Science. 2024. Vol. 1914. P. 147--161. DOI: 10.1007/978-3-031-52470-7{\_}13.

%2
\item \label{rfa:enlit:Barkalov2022}
Barkalov~K.A., et al. On solving the problem of finding kinetic parameters of catalytic isomerization of the pentane-hexane fraction using a parallel global search algorithm~// 
Mathematics. 2022. Vol. 10, No.~19. p. 3665. DOI: 10.3390/math10193665.

%8
\item \label{rfa:enlit:Gubaydullin2022}
Gubaydullin~I.M., Enikeeva~L.V., Barkalov~K.A., Lebedev~I.G., Silenko~D.G. Kinetic modeling of isobutane alkylation with mixed c4 olefins and sulfuric acid as a catalyst using the asynchronous global optimization algorithm~// Commun. Comput. Inf. Sci. 2022. Vol. 1618. P. 293--306. DOI: 10.1007/978-3-031-11623-0{\_}20.

%новые
\item \label{rfa:enlit:differential_evolution}
Storn~R., Price~K., Differential Evolution - a Simple and Efficient Heuristic for Global Optimization over Continuous Spaces // Journal of Global Optimization. 1997. Vol. 11. P. 341-359.

\item \label{rfa:enlit:dual_annealing}
Xiang~Y, Gubian~S, Suomela~B, Hoeng~J. Generalized Simulated Annealing for Efficient Global Optimization: the GenSA Package for R // The R Journal. 2013. Vol. 5, No. 1.

\item \label{rfa:enlit:direct}
Gablonsky~J., Kelley~C. A Locally-Biased form of the DIRECT Algorithm // Journal of Global Optimization. 2001. Vol. 21. P. 27-37.

\item \label{rfa:enlit:basinhopping}
Wales~D.J., Doye~J.P.K. Global Optimization by Basin-Hopping and the Lowest Energy Structures of Lennard-Jones Clusters Containing up to 110 Atoms // Journal of Physical Chemistry A. 1997. Vol. 101. P. 5111.

\item \label{rfa:enlit:shgo}
Endres~S.C., Sandrock~C., Focke~W.W. A simplicial homology algorithm for lipschitz optimisation // Journal of Global Optimization. 2018.

\item \label{rfa:enlit:Barkalov2021}
Barkalov K.A., Lebedev I.G., Gergel V.P. Parallel Global Search Algorithm with Local Tuning for Solving Mixed-Integer Global Optimization Problems // Lobachevskii Journal of Mathematics. Vol. 7. No. 42. 2021. P. 1492-1503.

\item \label{rfa:enlit:iOptPaper}
Sysoev~A.V., Kozinov~E.A., Barkalov~K.A., Lebedev~I.G., Karchkov~D.A., Rodionov~D.M. Frejmvork metodov intellektualnoj evristicheskoj optimizacii iOpt // V kn.: Superkompyuternye dni v Rossii: Trudy mezhdunarodnoj konferencii. 2023. S. 179-185.

\item \label{rfa:enlit:iOptDocs}
Dokumentaciya iOpt -- URL: https://iopt.readthedocs.io/ru/latest/ (data obrashheniya: 26.01.2025).

\item \label{rfa:enlit:iOptGithub}
Isxodnyj kod frejmvorka iOpt -- URL: https://github.com/aimclub/iOpt (data obrashheniya: 26.01.2025).

%6
\item \label{rfa:enlit:Gaviano2003}
Gaviano~M., Kvasov~D.E., Lera~D., Sergeyev~Y.D. Software for generation of classes of test functions with known local and global minima for global optimization~// ACM Trans. Math. Softw. 2003. Vol. 29, No.~4. P. 469--480.

\end{enumerate}

\begin{center}
	\textbf{АЛГОРИТМ РЕШЕНИЯ ЗАДАЧ ГЛОБАЛЬНОЙ ОПТИМИЗАЦИИ\break
    С ЧАСТИЧНО ОПРЕДЕЛЕННЫМИ ФУНКЦИЯМИ}
\end{center}

\begin{center}
	{М.А.~Усова, И.Г.~Лебедев, A.A.~Штанюк, К.А.~Баркалов }
\end{center}

\begin{center}
{ННГУ им. Н.И. Лобачевского, 603022, Нижний Новгород, Россия}
\end{center}

\noindent{УДК 519.853.4}
\newline
\begin{small}
% аннотация рус.
В статье обсуждается проблема поиска глобального минимума функции, которая задана в виде <<черного ящика>> и может быть не всюду определена в области поиска. Такие функции возникают при решении прикладных задач из-за особенностей оптимизируемого объекта или метода моделирования. Например, численная нестабильность метода при определенном сочетании параметров может привести к возникновению подобластей, в которых невозможно корректно провести моделирование и оценить значение целевой функции. В некоторых случаях эти подобласти известны, но в большинстве случаев информация о них отсутствует. Существование подобластей, в которых целевая функция является неопределенной, можно интерпретировать как наличие некоторых скрытых, заранее неизвестных ограничений задачи. Предлагаемый подход к решению такого рода задач является расширением информационно-статистического алгоритма глобального поиска и учитывает возможность неопределенных значений целевой функции в некоторых точках. 
В рамках предложенного алгоритма на каждом шаге поиска проводится оценка характеристик поисковых интервалов на основе значений целевой функции, вычисленных на их границах. 
В случае отсутствия информации о значениях функции для вычисления характеристики интервала используются оценка, учитывающая длину этого интервала. Для сокращения количества испытаний в подобластях, в которых функция не определена, предлагается использовать специальный параметр метода, позволяющий регулировать число точек испытаний в области невычислимости. В статье изложено описание модифицированного алгоритма глобального поиска для решения такого класса задач. Приведены результаты экспериментов на серии тестовых задач. Произведено сравнение с другими известными алгоритмами глобальной оптимизации. Приводится пример решения прикладной задачи настройки гиперпараметров метода машинного обучения, в которой настраиваемый метод работает некорректно при некоторых сочетаниях его параметров.

%Здесь текст аннотации, содержащий краткую постановку задачи и описание метода решения на русском языке объемом от 1000 знаков.

\textbf{Ключевые слова:} липшицева глобальная оптимизация, многоэкстремальные функции, функции вида <<черный ящик>>, частично определенные функции, скрытые ограничения, настройка гиперпараметров.
\end{small}

\section{Введение}
Глобальная оптимизация занимается разработкой методов, предназначенных для отыскания точек абсолютного (глобального) минимума многоэкстремальных функций. Возможность достоверной оценки глобального оптимума в таких задачах принципиально основана на наличии априорной информации о решаемой задаче, позволяющей связать возможные значения целевой функции с известными значениями в точках проведенных поисковых испытаний.

Весьма часто такая информация представляется в виде предположения, что целевая функция $\phi(y)$ удовлетворяет условию Липшица с неизвестной априори константой $L$. Это предположение можно интерпретировать (применительно к прикладным задачам) как отражение ограниченности мощностей, порождающих изменения в моделируемой системе. При этом целевая функция зачастую задаётся в виде <<черного ящика>>, является недифференцируемой и каждое вычисление ее значения в некоторой точке допустимой области может требовать значительных вычислительных ресурсов.

Для решения задач липшицевой глобальной оптимизации разработан ряд эффективных детерминированных методов 
[\ref{rfa:rulit:Jones2021},\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Birect2020},\ref{rfa:rulit:Sergeyev2017}]. Проведенные сравнения показывают, что детерминированные алгоритмы превосходят (по разным критериям) широко распространенные природно-естественные алгоритмы [\ref{rfa:rulit:Liberti2005},\ref{rfa:rulit:Sergeyev2018}].

Данная работа продолжает развитие одного из эффективных детерминированных методов решения задач липшицевой глобальной оптимизации -- ин\-фор\-ма\-ци\-он\-но-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}]. Данный алгоритм предполагает решение многомерных задач посредством их сведения к задачам одномерной оптимизации при помощи кривых, заполняющих пространство (кривых Пеано).

Отметим, что для оптимизации сложных объектов реального мира естественным является использование при расчетах сложных математических моделей, что, как следствие, существенно увеличивает трудоемкость поиска оптимума. За последние десятилетия специалистами в области оптимизации и параллельных вычислений было предложено множество способов снижения вычислительной сложности и ускорения используемых алгоритмов, связанных как с решением возникающих оптимизационных задач [\ref{rfa:rulit:Kvasov2013},\ref{rfa:rulit:Sergeyev2020}], так и с численным анализом исходных моделей [\ref{rfa:rulit:Dongarra2022},\ref{rfa:rulit:Duwe2020}].

Однако в последнее время актуальной становится принципиально новая проблема прикладных задач -- численная нестабильность исследуемых моделей в некоторых (заранее не известных) подобластях области изменения параметров. В таких подобластях невозможно корректно провести численное моделирование и вычислить значение целевой функции задачи. Данное явление можно интерпретировать или как наличие в задаче некоторых скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], или как наличие неизвестных областей, в которых целевая функция не является непрерывной [\ref{rfa:rulit:Audet2022}], или как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Sergeyev2003},\ref{rfa:rulit:Strongin2020}]. В такой постановке задача оптимизации существенно усложняется, т.к. область допустимых сочетаний параметров является заранее неопределенной.

%НОВОЕ
Характерным примером является задача настройки гиперпараметров методов искусственного интеллекта (ИИ) и машинного обучения (МО). Во многих случаях разница в эффективности (в заданной метрике) получаемого решения при разных значениях гиперпараметров может быть достаточно значительной. При этом вычислительная сложность настройки делает невозможным использование полного перебора. Данные особенности определили необходимость разработки различных методов интеллектуального автоматического подбора гиперпараметров. При этом проблема возникновения недопустимых комбинаций гиперпараметров, которая не имела особого значения в случае «ручной» настройки, для многих фреймворков, использующих методы интеллектуальной настройки, стала камнем преткновения. При этом сами настраиваемые методы могут вести себя абсолютно по-разному: в лучшем случае информируют пользователя о том, что задача не может быть решена, в худшем возвращают недопустимое значение (NaN, inf) в качестве найденного решения.

Отметим, что в распространенных фреймворках с открытым исходным кодом, предназначенных для решения задач оптимизации, отсутствует возможность работы с частично вычислимыми критериями.

В статье приведено описание алгоритма глобального поиска, адаптированного для работы с частично определенной целевой функцией. Данный алгоритм основан на подходе, предложенном авторами ранее в [\ref{rfa:rulit:Usova2024}]. Проведено исследование поведения других известных алгоритмов при решении задач с не всюду вычислимой целевой функцией. Продемонстрированы результаты численных экспериментов как с тестовыми задачами, так и с прикладной задачей настройки гиперпараметров метода LinearSVC при решении задачи классификации на известном наборе данных Iris.

\section{Постановка задачи}
В общем виде задача глобальной оптимизации может быть сформулирована следующим образом:
\begin{equation}\label{eq1} 
\phi^*=\phi(y^* )=\min_{y \in D}{\phi(y)},
\end{equation}
\[
D=\left\{ y \in R^N: a_i \leq y_i \leq b_i, \; 1 \leq i \leq N\right\},
\]
где $y=(y_1,y_2,...,y_N)$ -- вектор варьируемых параметров, $D$ -- $N$-мерный гиперкуб, $N$ -- размерность решаемой задачи.
О целевой функции $\phi (y)$ мы делаем следующие предположения.

\begin{enumerate}
\item{Целевая функция может быть многоэкстремальной, недифференцируемой и, более того, заданной в форме <<черного ящика>> (т.е. в виде некоторой подпрограммы, на вход которой подается аргумент, а выходом является соответствующее значение функции).}
\item{Каждое вычисление функции в некоторой точке допустимой области может требовать значительных вычислительных ресурсов.}
\item{Целевая функция удовлетворяет условию Липщица
\begin{equation}\label{eq3} 
| \phi (y')-\phi (y'') | \leq L \| y'-y'' \|, \; y',y'' \in D,
\end{equation}
где $0<L<\infty$ -- константа Липщица.}
\item{В некоторой подобласти $I$ области поиска $D$ (в частном случае, в одной или нескольких ее точках) целевая функция может быть не определена. Тогда функция $\phi(y)$ определена и вычислима лишь в подобласти $Q = D \backslash I$ (положительного объёма). Отметим, что исходя из опыта решения прикладных оптимизационных задач [\ref{rfa:rulit:Barkalov2022},\ref{rfa:rulit:Gubaydullin2022}], суммарный объем области невычислимости $I$ составляет небольшую долю объема области поиска $D$.}
\end{enumerate}

Последнее предположение делает невозможными применение известного ин\-фор\-ма\-ци\-он\-но-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска [\ref{rfa:rulit:Strongin2000}] или других методов липшицевой оптимизации [\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Sergeyev2017}]. Для решения таких задач нами была разработана модификация алгоритма глобального поиска (АГП-Н), основанная на одном из предложенных ранее подходов [\ref{rfa:rulit:Usova2024}].

\subsection{Редукция размерности}

Используя кривые типа развертки Пеано, однозначно отображающие отрезок $[0,1]$ на $N$-мерный единичный гиперкуб
\begin{equation}\label{eq2_} 
D=\left\{ y \in R^N: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\},
\end{equation}
исходную задачу (\ref{eq1}) можно редуцировать к одномерной задаче
\begin{equation}\label{eq2} 
f^*(x)=\phi(y(x^* ))=\min_{x \in [0,1]} \left\{ \phi(y(x)) \right\},
\end{equation}
что позволяет применить для ее решения эффективные алгоритмы одномерной оптимизации. В качестве иллюстрации на Рис.~\ref{fig_peano} представлены образы отрезка $[0,1]$ для $N=2$, полученные с помощью разверток с плотностью $p=4,5$.

\begin{figure}[h!]
	\center{\includegraphics[scale=0.70]{figures/peano.pdf}}
	\caption{Отображение отрезка $[0,1]$ с плотностью развертки: (а) $p=4$, (б) $p=5$}
	\label{fig_peano}
\end{figure}

Известно, что схема редукции размерности с использованием кривых Пеано сопоставляет многомерной задаче с липшицевой целевой функцией (\ref{eq1}) задачу (\ref{eq2}) с одномерной целевой функцией, удовлетворяющей условию Гёльдера
\begin{equation}\label{eq4} 
| f(x')-f(x'') | \leq K \rho(x',x''), \; x',x'' \in [0,1],
\end{equation}
где $\rho(x',x'') =  |x' - x''|^{1/N}$ -- метрика Гёльдера, $N$ -- размерность исходной задачи, а коэффициент $K$ связан с константой Липшица $L$ соотношением $K \leq 2L\sqrt {N+3}$ [\ref{rfa:rulit:Strongin2000}].

\section{Алгоритм решения задач с частично определенной целевой функцией}

Нашей целью является разработка алгоритма, в котором отсутствие значения целевой функции в некоторой точке не должно приводить к его некорректной работе. 
Для этого нужно предложить новые правила обработки поисковых интервалов, в граничных точках которых отсутствуют значения целевой функции. Например, расчет характеристики интервала с двумя невычислимыми или с невычислимой и граничной точками можно производить, основываясь на длине интервала. Расчет характеристики интервала с невычислимой и вычислимой точками можно производить по правилам работы с граничным интервалом.

Приведем подробное описание модификации алгоритма глобального поиска для не всюду вычислимой целевой функции (АГП-Н) для решения задач (\ref{eq2}).

\subsection{Описание алгоритма АГП-Н} \label{algorithm_discr}

На каждой итерации глобального поиска выполняется \textit{испытание}. Испытанием будем называть вычисление значения оптимизируемой функции $\phi (y(x))$ из (\ref{eq2}). Первое испытание будет осуществляться в серединной внутренней точке $x^1 \in (0,1)$. Выбор точки $x^{k+1}, k \geq 1$ очередного $(k+1)^\text{-го}$ испытания осуществляется на основе следующих правил.

\textit{Правило 1.} Перенумеровать (нижним индексом) точки $x^i, 0 \leq i \leq k$ предшествующих испытаний в порядке возрастания значений координаты, то есть
\begin{equation}\label{eq5} 
0=x_0 < x_1 < ... < x_i < ... < x_{k}=1
\end{equation}
и сопоставить им значения $z_i=f(x_i), 0 < i < k$, вычисленные в этих точках, и индекс $v_i=v(x_i)$, определяемый по правилу
\begin{equation}\label{eq6} 
v_i=v(x_i)=
  \begin{cases}
    -1, & {\quad \text{если } x_i \text{ -- граничная точка}},\\
    0, & {\quad \text{если } x_i \text{ -- невычислимая точка}},\\
    1, & {\quad \text{если } x_i \text{ -- внутренняя точка}}.
  \end{cases}
\end{equation}
Точки $x_0=0$ и $x_{k}=1$ введены дополнительно (значения $z_0$ и $z_{k}$ не определены).

\textit{Правило 2.} Вычислить текущую нижнюю оценку
\begin{equation}\label{eq7} 
\mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i},\quad v(x_{i-1})=v(x_i)=1, \quad 2 \leq i \leq k  \right\},\quad \Delta _i= (x_i-x_{i-1})^{1/N},
\end{equation}
константы Гёльдера $K$ из (\ref{eq4}) для редуцированной функции $f(x)$, где $r > 1$ -- параметр надёжности. Если $\mu = 0$, то принять $\mu = 1$.

\textit{Правило 3.} Определить текущее лучшее значение целевой функции
\begin{equation}\label{eq8} 
z^*= \min \left\{ f(x_i): v(x_i)=1, 0 < i < k \right\}.
\end{equation}

\textit{Правило 4.} Для каждого интервала $(x_{i-1},x_i),1 \leq i \leq k$, вычислить значение $R(i)$, называемое \textit{характеристикой} интервала, согласно выражению
\begin{equation}\label{eq9} 
R(i)=
  \begin{cases}
    \alpha{(1-\frac{1}{r})}^2 \Delta _i,& {\quad v(x_i)=v(x_{i-1})=0 \text{ или } v(x_i)=0 \text{ и } v(x_{i-1})=-1}\\
    {} & {\quad \text{или } v(x_i)=-1  \text{ и } v(x_{i-1})=0},\\
    \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1},\\
    2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad  v(x_{i-1})=-1 \text{ или } v(x_{i-1})=0, v(x_i)=1},\\
    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1 \text{ или } v(x_i)=0,}
  \end{cases}
\end{equation}
где $\alpha < 1.0$ - задаваемый параметр метода.

\textit{Правило 5.} Выбрать интервал $(x_{t-1},x_t)$ с максимальным значением характеристики $R(i)$: $R(t)= \max\{R(i): \; 1 \leq i \leq k\}$. Если максимальную характеристику имеют несколько интервалов, выбрать интервал %с большим количеством вычислимых точек
с минимальным номером.

\textit{Правило 6.} Провести очередное испытание в серединной точке интервала $(x_{t-1},x_t)$, если индексы его концевых точек не совпадают или оба соответствуют невычислимым точкам, т.е.
\begin{equation}\label{eq10} 
x^{k+1}=\frac {x_t+x_{t-1}}{2},\quad v(x_{t-1})\neq v(x_t) \text{ или } v(x_{t-1})=v(x_t)=0.
\end{equation}
В противном случае провести испытание в точке 
\begin{equation}\label{eq11} 
x^{k+1}= \frac {x_t+x_{t-1}}{2} -  \text{sign} {(z_t-z_{t-1})} \frac{1}{2r} \left[\frac {{|z_t-z_{t-1}|}}{\mu} \right]^N.
\end{equation}

\textit{Условие остановки.} Поиск завершен, если длина интервала $(x_{t-1},x_t)$, содержащего очередную точку $x^{k+1}$, не превышает заданной точности $\varepsilon$, то есть $\Delta _t \leq \varepsilon$, а одна из граничных точек данного интервала является внутренней точкой, т.е. $\min{\{v_{t-1}, v_t\}} = 1$, где $t$ из Правила 5, а $\varepsilon>0$ -- задаваемый параметр алгоритма.

%\begin{algorithm}
%\caption{Алгоритм глобального поиска}\label{alg:cap}
%\begin{algorithmic}
%\Require $\varphi(y), D, r, \epsilon, K_{max}$
%\Ensure $y_{min}$
%\State $k = 2$
%\State $t = 2$
%\State $\Omega_k = \left\{ (x_1 = 0, y_1 = y(x_1), z_1 = \varphi(y_1)), (x_2 = 1, y_2 = y(x_2), z_2 = \varphi(y_2)) \right\}$
%\While{ $\Delta_t \geq \epsilon \text{ and } k \leq K_{max}$}
%    \State Compute the estimate of the H\"older constant $\mu$ according to (\ref{eq7})
%    \For{$i=1$ \textbf{to} $k$}
%        \State Compute the characteristic $R(i)$ according to (\ref{eq9})
%    \EndFor
%    \State Determine the number $t$, for which $R(t) = \max \left\{ R(i), 1 \leq i \leq k \right\}$
%    \State Compute the new trial point $x^{k+1}$ according to (\ref{eq10}-\ref{eq11})
%    \State Perform a new trial at the point $x^{k+1}$
%    \State Add $(x^{k+1}, y^{k+1}, z^{k+1})$ to the set $\Omega_k$
%    \State $k=k + 1$
%\EndWhile
%\State $l = \arg \min \left\{ z_i, 1 < i \leq k \right\}$
%\State  $y_{min} = y_l$\\
%\Return $y_{min}$
%\end{algorithmic}
%\end{algorithm}

\textbf{Замечание 1 (о существовании вычислимого отрезка на кривой Пеано)}. В случае попадания точек испытаний в область невычислимости на первых итерациях, алгоритм будет выбирать лучший интервал на основе длин интервалов (т.е. будет строится равномерная сетка). Так как мы предполагаем, что область вычислимости целевой функции $Q = D \backslash I$ -- положительного объёма, то последующее сгущение сетки приведет к попаданию очередной точки испытания в область $Q$.

%\begin{figure}[h!]
%	\center{\includegraphics[scale=0.30]{figures/fig_0.png}}
%	\caption{Пример расстановки точек испытаний до момента попадания в вычислимую точку}
%	\label{fig_1}
%\end{figure}

\textbf{Замечание 2 (о параметре $\alpha$)}. Параметр $0 < \alpha \leq 1.0$ позволяет регулировать плотность сетки испытаний в области невычислимости $I$, $I \subset D$. Чем меньше параметр, тем менее перспективными для алгоритма будут интервалы с граничными точками в областях невычислимости. При максимально допустимом параметре $\alpha = 1.0$ метод густо усеивает точками испытаний подобласти, в которых функция не вычислима, возникает эффект <<черного пятна>> (точки испытаний в таких подобластях на рисунке обозначены черным цветом), см. Рис. \ref{task_61}a и Рис. \ref{task_61}б. При уменьшении параметра $\alpha$ точек этих испытаний становится значительно меньше, что наглядно видно на Рис. \ref{task_61}в и Рис. \ref{task_61}г.

\begin{figure}[h!]
	\center{\includegraphics[scale=0.15]{figures/task_61.png}}
	\caption{Распределение точек испытаний с различными видами невычислимых областей и настройкой параметра $\alpha$: (а) область невычислимости на границе области поиска, $\alpha = 1.0$; (б) случайно расположенные области невычислимости, $\alpha = 1.0$; (в) область невычислимости на границе области поиска, $\alpha = 0.08$; (г) случайно расположенные области невычислимости, $\alpha = 0.08$.}
	\label{task_61}
\end{figure}

\textbf{Замечание 3 (о связи с АГП)}. В случаях, когда задача не содержит скрытых ограничений или проводимые алгоритмом испытания не попадают в существующие области невычислимости, АГП-Н полностью повторяет работу своего прототипа -- алгоритма глобального поиска.

\textbf{Замечание 4 (о решении задач с дискретными параметрами)}. Алгоритм глобального поиска допускает обобщение для решения задач, в которых часть переменных являются непрерывными, а часть может принимать только дискретные значения [\ref{rfa:rulit:Barkalov2021}]. Это позволяет применять АГП-Н в задачах, где часть параметров объекта оптимизации может принадлежать некоторому дискретному множеству. Подобные задачи часто возникают при настройке гиперпараметров методов машинного обучения.

%\subsection{Схема распараллеливания}

%Пусть выполнено $n \geq 1$ итераций метода, в процессе которых были проведены испытания в $k(n)$ точках $x^i,1 \leq i \leq k$. Тогда точки $x^{k+1},…,x^{k+p}$ поисковых испытаний следующей $(n+1)$-ой итерации определяются в соответствии с правилами.

%Правила 1-4 параллельного алгоритма полностью повторяют правила 1-4 последовательного метода.

%\textit{Правило 5}. Характеристики $R(i), 1 \leq i \leq k+1$, упорядочить в порядке убывания
%\begin{equation}
%R(t_1) \geq R(t_2) \geq \dots \geq R(t_{k-1}) \geq R(t_{k+1})
%\end{equation}
%и выбрать $p$ наибольших характеристик с номерами интервалов $t_j, 1 \leq j \leq p$.

%\textit{Правило 6}. Провести новые испытания в точках $x^{k+j}, 1 \leq j \leq p$, вычисленных по формулам
%\begin{equation}
%x^{k+1}=\frac{x_{t_j}+x_{t_j-1}}{2},v(x_{t_j-1}) \neq v(x_{t_j}),
%\end{equation}
%\begin{equation}
%x^{k+1}=\frac{x_{t_j}+x_{t_j-1}}{2} -sign(z_{t_j}-z_{t_j-1}) \frac{1}{2r_v} {\left[\frac{|z_{t_j}-z_{t_j-1}|}{\mu_v}\right]}^N, v=v(x_{t_j-1})=v(x_{t_j}).
%\end{equation}

%Алгоритм прекращает работу, если выполняется условие $\Delta_{t_j} \leq \varepsilon$ хотя бы для одного номера $t_j, 1 \leq j \leq p$, при этом одна из граничных точек, соответствующего этому номеру интервала, является внутренней точкой, т.е. $min{\{v_{t_j-1}, v_t_j\}} = 1$; здесь $\varepsilon>0$ есть заданная точность.

\section{Программная реализация} \label{iOpt_discr}

Алгоритм глобального поиска, предназначенный для решения задач с не всюду вычислимой целевой функцией (АГП-Н), был реализован на базе фреймворка методов интеллектуальной оптимизации iOpt [\ref{rfa:rulit:iOptPaper}] с открытым исходным кодом [\ref{rfa:rulit:iOptGithub}].

Фреймворк позволяет проводить точную настройку параметров моделей и методов, используемых в прикладных исследованиях в различных научных областях. Характерными примерами задач, решаемых фреймворком iOpt, являются задачи настройки гиперпараметров методов машинного обучения, а также методов эвристической оптимизации. Область применения фреймворка фокусируется на промышленных задачах и методах. В таких задачах часто возникает рассматриваемая в данном исследовании проблема частичной определенности критериев в отсутствии формульного описания исследуемой модели (модель вида «черный ящик»).

На Рис.~\ref{fig_iOpt} приведена блок-схема общего алгоритма функционирования фреймворка.

\begin{figure}[h!]
	\center{\includegraphics[scale=0.68]{figures/iOpt.pdf}}
	\caption{Блок-схема общего алгоритма функционирования фреймворка iOpt}
	\label{fig_iOpt}
\end{figure}

Реализация фреймворка iOpt выполнена с использованием объектно-ориентированного программирования в виде системы взаимосвязанных классов. Реализация описанного в разделе \ref{algorithm_discr} алгоритма содержится в базовом классе Method. Описанный далее в разделе \ref{GKLS_HC} генератор задач GKLS со скрытыми ограничениями (GKLS-HC) реализован в фреймворке iOpt в виде класса GKLSHiddenConstraint -- наследника интерфейса Problem. В фреймворке добавлена возможность визуализации процесса решения задач со скрытыми ограничениями с использованием StaticPainterNDListener. Доступны различные способы отрисовки линий уровня целевой функции: по равномерной сетке, только по точкам поисковых испытаний, по точкам испытаний с использованием интерполяции. Подробное описание возможностей фреймворка представлено в документации [\ref{rfa:rulit:iOptDocs}].

\section{Результаты вычислительных экспериментов}

Решение серии задач с известными глобальными оптимумами является одним из традиционных способов оценки качества работы методов глобального поиска. 
Однако известные наборы тестовых задач не содержат областей невычислимости целевой функции, поэтому для тестирования разработанного алгоритма АГП-Н был предложен способ генерации тестовых задач со скрытыми ограничениями. В данном разделе дано краткое описание тестовых задач, а также приведены результаты экспериментов, демонстрирующие надёжность глобального поиска и эффективность разработанного алгоритма оптимизации.

\subsection{Описание тестовых задач} \label{GKLS_HC}

Для создания тестовых задач GKLS-HC использовался генератор GKLS [\ref{rfa:rulit:Gaviano2003}], позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами.

Генерация задачи состоит в определении выпуклой квадратичной функции, дополненной полиномами более высокого порядка для введения локальных минимумов. Каждый тестовый класс, предоставляемый генератором GKLS, состоит из 100 функций, построенных случайным образом, и определяется размером области поиска, количеством локальных минимумов, значением глобального минимума, радиусом области притяжения глобального минимума, расстоянием от глобального минимума до вершины квадратичной функции. Другие необходимые параметры, такие как точки всех локальных минимумов, их области притяжения и значения, выбираются генератором случайным образом.

Данные функции были дополнены областями невычислимости в форме эллипсоидов (координаты центров и длины радиусов генерировались случайно), при попадании в которые функция возвращает исключение, сообщающее о невозможности вычислить ее значение в заданной точке.

В текущей версии фреймворк iOpt использует генератор GKLS для формирования задач класса Simple с размерностью от 2 до 5, числом локальных минимумов равным 10 и областью поиска от $-1$ до $1$ по каждой переменной. Радиус области притяжения глобального минимума и расстояние от глобального минимума до вершины квадратичной функции варьируется в зависимости от размерности задачи. Количество генерируемых областей невычислимости задавалось равным 4. Области невычислимости могут пересекаться, при этом не включают точку известного глобального минимума функций GKLS по построению. Радиусы областей невычислимости выбираются случайно из диапазона $[0.05, 0.25]$.

Если говорить о программной реализации, то генератор задач реализован в виде класса GKLSHiddenConstraint, конструктор которого принимает размерность генерируемой задачи (от 2 до 5), номер функции (от 1 до 100), а также параметр для настройки генератора случайных чисел для воспроизведения экспериментов, после чего автоматически устанавливает остальные параметры генератора.

\subsection{Решение тестовых задач с помощью АГП-Н}

Вычислительный эксперимент был проведен с использованием фреймворка iOpt, описанного в разделе \ref{iOpt_discr}. АГП-Н запускался на серии задач GKLS-HC с точностью поиска минимума $\varepsilon = 0.001$, параметром надёжности $r = 4.2$ и двумя значениями параметра $\alpha$: $\alpha = 0.08$ и $\alpha = 0.008$. Для полноценного сравнения алгоритм АГП-Н также был запущен на серии задач GKSL (т.е. на задачах без подобластей, в которых целевая функция является неопределенной). Результаты эксперимента приведены в таблице \ref{iOpt_gkls}.

\begin{table}[h!]
\centering
\caption{Результаты решения задач GKLS и GKLS-HC средствами фреймворка iOpt}
\begin{tabular}{||r|c|c|c||}
\hline
\multicolumn{1}{||c|}{} & alpha & \begin{tabular}[c]{@{}c@{}}среднее\\ число\\ итераций\end{tabular} & \begin{tabular}[c]{@{}c@{}}кол-во\\ решенных\\ задач\end{tabular} \\ \hline
\multirow{2}{*}{GKLS-HC} & 0.08  & 2151 & 100 / 100 \\ \cline{2-4} 
                                        & 0.008 & 1635 & 100 / 100 \\ \hline
GKLS                                    & -     & 1510 & 100 / 100 \\ \hline
\end{tabular}
\label{iOpt_gkls}
\end{table}

На Рис. \ref{oper_charact} показаны операционные характеристики, которые указывают для каждого количества итераций число задач из тестовой выборки, решенное с заданной точностью $\varepsilon = 0.001$.

\begin{figure}[h!]
	\center{\includegraphics[scale=0.75]{figures/oper_charact_transp.pdf}}
	\caption{Операционные характеристики алгоритмов АГП-Н и АГП}
	\label{oper_charact}
\end{figure}

Результаты проведенных экспериментов показывают, что после выполнения 2000 итераций АГП-Н с параметром $\alpha = 0.008$ будет получена оценка оптимума с требуемой точностью для 80\% задач. Для решения оставшихся 20\% задач требуется (в худшем случае) 4315 поисковых испытаний. Отметим, что за счёт регулировки параметра $\alpha$ получилось добиться такой же скорости работы, как у  АГП при решении серии задач GKLS (в терминах числа испытаний, требующихся для решения задач с заданной точностью).

Пример визуализации процесса решения задачи GKLS-HC №~38 представлен на Рис. \ref{iOpt_result}. Черными точками отмечены точки испытаний, попавшие в подобласти, в которых функция не определена. Красной точкой отмечено найденное алгоритмом решение.

\begin{figure}[h!]
	\center{\includegraphics[scale=1.1]{figures/iOpt_result.pdf}}
	\caption{Графики линий уровня задачи GKLS-HC №~38 с точками проводимых испытаний, полученные средствами фреймворка iOpt: (а) $\alpha = 0.08$, (б) $\alpha = 0.008$ }
	\label{iOpt_result}
\end{figure}

\subsection{Сравнение АГП-Н с известными алгоритмами оптимизации}

\subsubsection{Описание алгоритмов оптимизации}

В данной статье сравнение проводилось с алгоритмами глобальной оптимизации библиотеки \textit{scipy.optimize}, а именно с методами \textit{differential\_evolution} [\ref{rfa:rulit:differential_evolution}], \textit{dual\_annealing} [\ref{rfa:rulit:dual_annealing}], \textit{direct} [\ref{rfa:rulit:direct}], \textit{basinhopping} [\ref{rfa:rulit:basinhopping}], \textit{shgo} [\ref{rfa:rulit:shgo}], \textit{brute}.

Метод дифференциальной эволюции (differential evolution) -- один из самых известных генетических алгоритмов вещественной оптимизации. Во многих практических задачах метод обеспечивает приемлемое решение, т.к. не использует информацию о градиенте функции, а схемы мутации и рекомбинации позволяют ему <<выходить>> из локальных минимумов.

Двойной отжиг (dual annealing) -- алгоритм глобальной оптимизации, основанный на комбинировании алгоритмов классического и быстрого имитационного отжига. Алгоритм имитирует процесс отжига в металлургии. Отжиг заключается в нагревании и контролируемом охлаждении металла для формирования устойчивой кристаллической решётки: при высокой температуре системы алгоритм использует более широкий диапазон решений, при снижении температуры диапазон поиска становится всё меньше, пока не будет найден глобальный оптимум.

Direct (DIviding RECTangles), как и АГП, предназначен для решения задач липшицевой глобальной оптимизации. Особенностью метода является отказ от оценивания константы Липшица. В процессе своей работы алгоритм разбивает область поиска на гиперинтервалы, среди которых выбираются потенциально оптимальные (подлежащие дальнейшему разбиению), после чего в их центрах проводится вычисление значений целевой функции.

Перепрыгивание через бассейн (basin-hopping) -- алгоритм глобальной оптимизации, осуществляющий поиск оптимума путём последовательного выполнения локальной оптимизации (по умолчанию используя метод Монте-Карло) с последующим принятием или отклонением новых координат на основе текущего лучшего значения функции.

SHGO (simplicial homology global optimization) сравнительно новый алгоритм глобальной оптимизации, основанный на применении симплициальной гомологии. Алгоритм использует концепции из комбинаторной интегральной теории гомологии для поиска подобластей, которые приблизительно локально выпуклы. В отличие от многих других алгоритмов глобальной оптимизации, использующих теорию графов и методы кластеризации, SHGO создает начальные точки, которые сходятся к уникальным локальным минимумам, что улучшает его производительность.

Метод brute реализует метод <<грубой силы>> (brute force), т.е. полный перебор.

Алгоритмы \textit{differential\_evolution}, \textit{dual\_annealing}, \textit{basinhopping} относятся к классу стохастических алгоритмов (то есть работают с использованием случайных чисел и каждый запуск поиска может найти другое решение), потому поиск глобального минимума с использованием данных алгоритмов требует многократных запусков.

\subsubsection{Результаты проведенного сравнения}

В рамках экспериментов данного исследования были получены следующие результаты.

При наличии у задачи скрытых ограничений худшее поведение демонстрируют алгоритмы differential\_evolution и brute. Алгоритм differential\_evolution полностью исчерпывает все выделенные ему итерации, не находя при этом корректный ответ. Полный перебор в свою очередь в качестве результата работы может вернуть что угодно.

Алгоритмы dual\_annealing и basinhopping в ряде случаев справляются с поставленной оптимизационной задачей. При этом в рамках эксперимента количество вычислений целевой функции алгоритмом dual\_annealing составило порядка 4000, в то время как basinhopping мог выполнять и 140000 вычислений для одной задачи.

Алгоритмы direct и shgo демонстрируют хорошие результаты, хоть и не всегда обладают сходимостью. Среднее число итераций алгоритма direct составило 94 (соответствует 1981 вычислениям целевой функции). Алгоритм shgo в каждой задаче полностью исчерпал выделенные ему 200 итерации, выполняя в среднем 4482 вычислений целевой функции. В сравнении с данными методами АГП-Н работает эффективнее (отметим, что одна итерация АГП-Н соответствует одному вычислению значения целевой функции) и обладает лучшей сходимостью к глобальному оптимуму.

\begin{table}[h!]
\centering
\caption{Возможные выявленные результаты работы алгоритмов глобальной оптимизации библиотеки \textit{scipy.optimize} на задачах GKLS-HC}
\begin{tabular}{||r|c|c|c|c||}
\hline
\multirow{2}{*}{алгоритм} & \multicolumn{4}{c||}{возможный результат поиска} \\ \cline{2-5} 
 & \multicolumn{1}{c|}{глобальный минимум} & \multicolumn{1}{c|}{локальный минимум} & \multicolumn{1}{c|}{nan} & \multicolumn{1}{c||}{1e+100} \\ \hline
АГП-Н                   &     \checkmark     &                   &              &             \\ \hline
differential\_evolution &                    &                   &  \checkmark  &             \\ \hline
dual\_annealing         &     \checkmark     &                   &  \checkmark  &             \\ \hline
basinhopping            &     \checkmark     &    \checkmark     &              &  \checkmark \\ \hline
direct                  &     \checkmark     &    \checkmark     &              &             \\ \hline
shgo                    &     \checkmark     &    \checkmark     &              &             \\ \hline
brute                   &     \checkmark     &    \checkmark     &  \checkmark  &  \checkmark \\ \hline
\end{tabular}
\label{comparision}
\end{table}

В таблице \ref{comparision} приведена информация о возможном поведении исследованных алгоритмов с задачами с не всюду определенной целевой функцией, обнаруженная при решении задач GKLS-HC. Алгоритм АГП-Н всегда находит глобальный минимум, тогда как differential\_evolution всегда возвращает значение nan. Остальные алгоритмы могут вернуть как глобальный, так и локальный минимум, при этом алгоритмы dual\_annealing и basinhopping в ряде случаев могут вернуть значения nan и 1e+100 соответственно, а алгоритм brute, как уже было отмечено, и вовсе может вернуть в качестве ответа что угодно.

\subsection{Решение задачи настройки гиперпараметров метода LinearSVC}

В качестве модельной задачи оптимизации, в которой встречается недопустимое сочетание параметров, рассмотрим настройку гиперпараметров метода LinearSVC из библиотеки scikit-learn при решении задачи классификации. Исследование проводилось на классическом датасете Iris, который содержит характеристики трех различных видов ириса и часто используется в качестве тестового примера для алгоритмов машинного обучения. Набор данных Iris состоит из 150 записей (по 50 записей на каждый вид ириса), содержащих по пять атрибутов. Датасет включен в библиотеку машинного обучения scikit-learn.

Решалась задача максимизации метрики $Macro F1$ (характеризующей качество построения классификатора) в зависимости от трех гиперпараметров:
\begin{itemize}
\item вещественного коэффициента регуляризации $C$, $C \in [1,6]$;
\item дискретного типа функции потерь $loss$, $loss \in \{hinge, squared\_hinge\}$;
\item дискретного параметра $dual$, $dual \in \{true, false\}$, который позволяет выбирать между решением двойственной или основной задачи оптимизации.
\end{itemize}

Для возможности воспроизведения результатов параметр $random\_state$ был зафиксирован в значении 10. Остальные параметры метода LinearSVC рассматривались заданными по умолчанию. Максимизация метрики $F1$ проводилась с использованием нескольких алгоритмов:
\begin{itemize}
\item GridSearchCV -- метод полного перебора по равномерной сетке в области изменения гиперпараметров;
\item  RandomizedSearchCV -- метод случайного поиска в области изменения гиперпараметров;
\item  АГП-Н, предложенный в данной статье; в алгоритме использовался параметр $r=3.5$ и было задано ограничение на число испытаний $K_{max}=100$.
\end{itemize}

Отметим, что сочетание гиперпараметров $loss = hingle$ и $dual = False$ приводит к «падению» метода LinearSVC (метод выдает выбрасывает исключение и не выдает значения целевой метрики). В процессе своей работы оба алгоритма из scikit-learn решили задачу с выдачей предупреждения о возникновении недопустимого сочетания параметров; было найдено значение целевой метрики $F1 = 0.96$.
Алгоритм АГП-Н также решил поставленную задачу и нашел лучшее значение $F1 = 0.973$ при гиперпараметрах $C=1.3125$, $loss=squared\_hinge$, $dual = False$.

\section{Заключение}

В данной работе исследовалась проблема поиска глобального минимума вычислительно трудоемкой функции вида <<черный ящик>>. По сравнению с традиционной формулировкой, используемой в глобальной оптимизации, рассмотренная задача имеет важное отличие: предполагается, что целевая функция может быть не определена в некоторых подобластях области поиска (в частном случае, в одной или нескольких ее точках). В таких подобластях невозможно корректно провести численное моделирование и вычислить значение целевой функции.

Основная трудность рассматриваемых задач заключается в невозможности заранее узнать о подобластях невычислимости; фактически, область допустимых сочетаний параметров является заранее неопределенной. Данное явление можно интерпретировать как наличие в задаче некоторых скрытых ограничений.

В статье изложено подробное описание модифицированного алгоритма глобального поиска АГП-Н для решения исследуемого класса задач и генератора тестовых функций GKLS-HC, основанного на известном генераторе GKLS. Приведенные в статье результаты экспериментов, полученные при решении серии тестовых задач, демонстрируют надёжность разработанного алгоритма. Результаты работы АГП-Н сравнивались с результатами, полученными с помощью стандартных алгоритмов оптимизации из библиотеки \textit{scipy.optimize}. В рамках данного эксперимента разработанный алгоритм показал свою эффективность.

Также с помощью АГП-Н была решена задача настройки гиперпараметров алгоритма классификации LinearSVC. Известно, что сочетание некоторых гиперпараметров LinearSVC приводит к <<падению>> метода. Разработанный в ходе исследования и реализованный в библиотеке iOpt метод АГП-Н успешно решил поставленную задачу, найдя лучшее сочетание гиперпараметров по сравнению со стандартными алгоритмами из библиотеки scikit-learn.

%В статье было приведено описание одной из исследованных ранее модификаций алгоритма глобального поиска для решения такого класса задач. Приведены теоретические обоснования сходимости алгоритма. Теоретические результаты полностью соответствуют картине, полученной в ходе экспериментальных исследований эффективности алгоритма. Разработанная модификация обладает недостатком в виде избыточных испытаний в области невычислимости (в силу порождения алгоритмом предельных точек в данных областях). Данного явления можно избежать с использованием более сложной модификации, восстанавливающей значения в невычислимых точках, основываясь на предположении о липшицевости целевой функции (что было отражено в экспериментальных исследованиях). Теоретическое обоснование данной модификации является объектом будущих исследований.

\textbf{Благодарности.} Работа выполнена при поддержке Министерства науки и высшего образования РФ (проект № FSWR-2023-0034) и научно-образовательного математического центра <<Математика технологий будущего>>.

\section*{Список литературы}
% оформленный в соответствии с требованиями ГОСТ
\begin{enumerate}
%7
%\item \label{rfa:rulit:Grishagin2016_2}
%Grishagin~V.A., Israfilov~R.A. Global search acceleration in the nested optimization scheme~// AIP Conference Proceedings. 2016. Vol. 1738. p. 400010. DOI: 10.1063/1.4952198.

%9
\item \label{rfa:rulit:Jones2021}
Jones~D., Martins~J. The direct algorithm: 25 years later~// J. Glob. Optim. 2021. Vol. 79, No,~3. P. 521--566. DOI: 10.1007/s10898-020-00952-6.

%12
\item \label{rfa:rulit:PaulaviciusZilinskas2014}
Paulavi{\v c}ius~R. and {\v Z}ilinskas~J. Simplicial Global Optimization. New York: Springer, 2014. DOI: 10.1007/978-1-4614-9093-7.

%13
\item \label{rfa:rulit:Birect2020}
Paulavi{\v c}ius~R., Sergeyev~Y.D., Kvasov~D.E., {\v Z}ilinskas~J. Globally-biased {BIRECT} algorithm with local accelerators for expensive global optimization~// 
Expert Syst. Appl. 2020. Vol. 144. p. 113052. DOI: 10.1016/j.eswa.2019.113052.

%14
\item \label{rfa:rulit:Sergeyev2017}
Сергеев~Я.Д., Квасов~Д.Е. Диагональные методы глобальной оптимизации.  М.: Физматлит, 2008.

%11
\item \label{rfa:rulit:Liberti2005}
Liberti~L., Kucherenko~S. Comparison of deterministic and stochastic approaches to global optimization~// Int. Trans. Oper. Res. 2005. Vol. 12. P. 263--285.

%15
\item \label{rfa:rulit:Sergeyev2018}
Sergeyev~Y.D., Kvasov~D.E., Mukhametzhanov~M.S. On the efficiency of nature-inspired metaheuristics in expensive global optimization with limited budget~// Sci. Rep. 2018. Vol. 8, No.~1. p. 435.

%16
\item \label{rfa:rulit:Sergeyev2013}
Sergeyev~Y.D., Strongin~R.G., Lera~D. Introduction to Global Optimization Exploiting Space-Filling Curves. New York: Springer Briefs in Optimization, 2013. DOI: 10.1007/978-1-4614-8042-6.

%20
\item \label{rfa:rulit:Strongin2000}
Strongin~R.G., Sergeyev~Y.D. Global optimization with non-convex constraints. Sequential and parallel algorithms. Dordrecht: Kluwer Academic Publishers, 2000.

%10
\item \label{rfa:rulit:Kvasov2013}
Квасов~Д.Е., Сергеев~Я.Д. Методы липшицевой глобальной оптимизации в задачах управления~// Автоматика и телемеханика. 2013. № 9. C. 3--19.

%17
\item \label{rfa:rulit:Sergeyev2020}
Sergeyev~Y.D., Candelieri~A., Kvasov~D.E., Perego~R. Safe global optimization of expensive noisy black-box functions in the $\delta$-Lipschitz framework~// 
Soft Comput. 2020. Vol. 24, No.~23. P. 17715--17735. DOI: 10.1007/s00500-020-05030-3.

%4
\item \label{rfa:rulit:Dongarra2022}
Dongarra~J.J. The evolution of mathematical software~// Commun. ACM. 2022. Vol. 65, No.~12. P. 66--72. DOI: 10.1145/3554977.

%5
\item \label{rfa:rulit:Duwe2020}
Duwe~K., et al. State of the art and future trends in data reduction for high-performance computing~// Supercomput. Front. Innov. 2020. Vol. 7, No.~1. DOI: 10.14529/jsfi200101.

%19
\item \label{rfa:rulit:Stripinis2021}
Stripinis~L., Paulavi{\v c}ius~R. A new {DIRECT}-{GLh} algorithm for global optimization with hidden constraints~// Optim. Lett. 2021. Vol. 15, No.~6. P. 1865--1884.
DOI: 10.1007/s11590-021-01726-z.

%1
\item \label{rfa:rulit:Audet2022}
Audet~C., Batailly~A., Kojtych~S. Escaping unknown discontinuous regions in blackbox optimization~// SIAM J. Optim. 2022. Vol. 32, No.~3. P. 1843--1870. DOI: 10.1137/21m1420915.

%3
\item \label{rfa:rulit:Candelieri2019}
Candelieri~A. Sequential model based optimization of partially defined functions
under unknown constraints~// J. Glob. Optim. 2019. Vol. 79, No.~2. P. 281--303. DOI: 10.1007/s10898-019-00860-4.

%18
\item \label{rfa:rulit:Sergeyev2003}
Баркалов~К.А., Стронгин~Р.Г. Метод глобальной оптимизации с адаптивным порядком проверки ограничений~// Журн. вычисл. матем. и матем. физ. 2002. Т.~42. №~9. С.~1338--1350.

%21
\item \label{rfa:rulit:Strongin2020}
Strongin~R.G., Barkalov~K.A., Bevzuk~S.A. Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints~// 
Soft Comput. 2020. Vol. 24, No.~16. P. 11853--11865. DOI: 10.1007/s00500-020-05078-1.

\item \label{rfa:rulit:Usova2024}
Usova~M.A., Barkalov~K.A. An Algorithm for Finding the Global Extremum of a Partially Defined Function~// Communications in Computer and Information Science. 2024. Vol. 1914. P. 147--161. DOI: 10.1007/978-3-031-52470-7{\_}13.

%2
\item \label{rfa:rulit:Barkalov2022}
Barkalov~K.A., et al. On solving the problem of finding kinetic parameters of catalytic isomerization of the pentane-hexane fraction using a parallel global search algorithm~// 
Mathematics. 2022. Vol. 10, No.~19. p. 3665. DOI: 10.3390/math10193665.

%8
\item \label{rfa:rulit:Gubaydullin2022}
Gubaydullin~I.M., Enikeeva~L.V., Barkalov~K.A., Lebedev~I.G., Silenko~D.G. Kinetic modeling of isobutane alkylation with mixed c4 olefins and sulfuric acid as a catalyst using the asynchronous global optimization algorithm~// Commun. Comput. Inf. Sci. 2022. Vol. 1618. P. 293--306. DOI: 10.1007/978-3-031-11623-0{\_}20.

%новые
\item \label{rfa:rulit:differential_evolution}
Storn~R., Price~K., Differential Evolution - a Simple and Efficient Heuristic for Global Optimization over Continuous Spaces // Journal of Global Optimization. 1997. Vol. 11. P. 341-359.

\item \label{rfa:rulit:dual_annealing}
Xiang~Y, Gubian~S, Suomela~B, Hoeng~J. Generalized Simulated Annealing for Efficient Global Optimization: the GenSA Package for R // The R Journal. 2013. Vol. 5, No. 1.

\item \label{rfa:rulit:direct}
Gablonsky~J., Kelley~C. A Locally-Biased form of the DIRECT Algorithm // Journal of Global Optimization. 2001. Vol. 21. P. 27-37.

\item \label{rfa:rulit:basinhopping}
Wales~D.J., Doye~J.P.K. Global Optimization by Basin-Hopping and the Lowest Energy Structures of Lennard-Jones Clusters Containing up to 110 Atoms // Journal of Physical Chemistry A. 1997. Vol. 101. P. 5111.

\item \label{rfa:rulit:shgo}
Endres~S.C., Sandrock~C., Focke~W.W. A simplicial homology algorithm for lipschitz optimisation // Journal of Global Optimization. 2018.

\item \label{rfa:rulit:Barkalov2021}
Barkalov K.A., Lebedev I.G., Gergel V.P. Parallel Global Search Algorithm with Local Tuning for Solving Mixed-Integer Global Optimization Problems // Lobachevskii Journal of Mathematics. Vol. 7. No. 42. 2021. P. 1492-1503.

\item \label{rfa:rulit:iOptPaper}
Сысоев~А.В., Козинов~Е.А., Баркалов~К.А., Лебедев~И.Г., Карчков~Д.А., Родионов~Д.М. Фреймворк методов интеллектуальной эвристической оптимизации iOpt // В кн.: Суперкомпьютерные дни в России: Труды международной конференции. 2023. С. 179-185.

\item \label{rfa:rulit:iOptGithub}
Исходный код фреймворка iOpt -- URL: https://github.com/aimclub/iOpt (дата обращения: 26.01.2025).

\item \label{rfa:rulit:iOptDocs}
Документация iOpt -- URL: https://iopt.readthedocs.io/ru/latest/ (дата обращения: 26.01.2025).

%6
\item \label{rfa:rulit:Gaviano2003}
Gaviano~M., Kvasov~D.E., Lera~D., Sergeyev~Y.D. Software for generation of classes of test functions with known local and global minima for global optimization~// ACM Trans. Math. Softw. 2003. Vol. 29, No.~4. P. 469--480.

\end{enumerate}
%\newline

\section*{Краткая биография}
\paragraph{Усова Марина Андреевна} -- ассистент кафедры Математического обеспечения и суперкомпьютерных технологий Института информационных технологий, математики и механики ННГУ им. Н.И. Лобачевского. Область научных интересов: модели и методы решения задач глобальной оптимизации, разработка средств визуализация данных. E-mail: usova@itmm.unn.ru. Адрес: 603022, г. Нижний Новгород, пр. Гагарина, 23.
\paragraph{Лебедев Илья Генадьевич} -- заведующий лабораторией суперкомпьютерных технологий и высокопроизводительных вычислений кафедры Математического обеспечения и суперкомпьютерных технологий Института информационных технологий, математики и механики национального исследовательского нижегородского государственного университета им. Н.И. Лобачевского, e-mail: ilya.lebedev@itmm.unn.ru. Область научных интересов: методы глобальной и локальной оптимизации, параллельные алгоритмы, программирование для графических процессоров, CUDA. Является автором и соавтором 61 научной работы.
\paragraph{Штанюк Антон Александрович} -- кандидат технических наук, доцент, доцент кафедры информатики и автоматизации научных исследований национального исследовательского нижегородского государственного университета им. Н.И. Лобачевского, e-mail:anton.shtanyuk@itmm.unn.ru. Количество печатных работ 53. Область научных интересов: моделирование и проектирование сложных систем, языки и технология программирования, обучение программированию, алгоритмы и структуры данных.
\paragraph{Баркалов Константин Александрович} -- 
доктор технических наук, доцент, заведующий кафедры Математического обеспечения и суперкомпьютерных технологий Института информационных технологий, математики и механики ННГУ им. Н.И. Лобачевского. Область научных интересов: математические модели, методы и программные средства для решения задач глобальной оптимизации; параллельные вычисления. Является автором и соавтором более 140 научных работ в указанных областях. E-mail: konstantin.barkalov@itmm.unn.ru.

\section*{Biography}

\paragraph{Marina Usova} -- Assistant of the Department of Mathematical Software and Supercomputer Technologies of the Institute of Information Technologies, Mathematics and Mechanics, Lobachevsky State University of Nizhny Novgorod. Research interests: models and methods for solving global optimization problems, development of data visualization tools. E-mail: usova@itmm.unn.ru. Address: 603022, Nizhny Novgorod, Gagarin Ave., 23.
\paragraph{Ilya Lebedev} -- PhD., Head of the Laboratory of Supercomputer Technologies and High-Performance Computing, Department of Mathematical Software and Supercomputing Technologies, Institute of Information Technology, Mathematics and Mechanics, Nizhny Novgorod State University. N.I. Lobachevsky. e-mail: ilya.lebedev@itmm.unn.ru. His research interests include algorithms of global and local optimization, parallel algorithms, GPU programming, CUDA. He is the author or coauthor more than 30 papers in these areas.
\paragraph{Anton Shtanyuk} -- PhD., Associate Professor at the Department of Informatics and Research Automation of the Institute of Information Technology, mathematics and mechanics at Lobachevsky State University of Nizhny Novgorod, mail:anton.shtanyuk@itmm.unn.ru The number of publications: 50. Research interests: modelling and design of complex systems, programming languages and technology, programming teaching, algorithms and data structures.
\paragraph{Konstantin Barkalov} -- Doctor of Technical Sciences, Professor, Head of the Department of Mathematical Software and Supercomputer Technologies, Institute of Information Technologies, Mathematics and Mechanics, Lobachevsky State University of Nizhny Novgorod. Research interests: mathematical models, methods and software for solving global optimization problems; parallel computing. E-mail: konstantin.barkalov@itmm.unn.ru.

\end{document}

### 2025-02-04 2025/NumMet/NMP_authorGuidelinesRu_2023.tex
added template of paper with proof of convergence

--- 2024-09-29 2024/NumMet/NMP_authorGuidelinesRu_2023.tex
@@
 где $y=(y_1,y_2,...,y_N)$ -– вектор варьируемых параметров, $D$ -- $N$-мерный гиперкуб, $N$ –- размерность решаемой задачи.
-Используя кривые типа развертки Пеано, однозначно отображающие отрезок $[0,1]$ на $N$-мерный единичный гиперкуб
-\begin{equation}\label{eq2_} 
-D=\left\{ y \in R^N: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\},
-\end{equation}
-исходную задачу (\ref{eq1}) можно редуцировать к одномерной задаче
-\begin{equation}\label{eq2} 
-f^*(x)=\phi(y(x^* ))=\min_{x \in [0,1]} \left\{ \phi(y(x)) \right\},
-\end{equation}
-что позволяет применить для ее решения эффективные алгоритмы одномерной оптимизации. О целевой функции $\phi (y)$ мы делаем следующие предположения.
+О целевой функции $\phi (y)$ мы делаем следующие предположения.
 
@@
 \end{equation}
-где $L$ -- константа Липщица.
-Известно, что схема редукции размерности с использованием кривых Пеано сопоставляет многомерной задаче с липшицевой целевой функцией (\ref{eq1}) задачу (\ref{eq2}) с одномерной целевой функцией, удовлетворяющей условию Гёльдера
-\begin{equation}\label{eq4} 
-| f(x')-f(x'') | \leq K \rho(x',x''), \; x',x'' \in [0,1],
-\end{equation}
-где $\rho(x',x'') =  |x' - x''|^{1/N}$ -- метрика Гёльдера, $N$ -- размерность исходной многомерной задачи, а коэффициент $K$ связан с константой Липшица $L$ соотношением $K \leq 2L\sqrt {N+3}$ [\ref{rfa:rulit:Strongin2000}].}
-\item{В некоторых подобластях $I_1, I_2, ..., I_k: I_1 \cup I_2 \cup ...
-\cup I_k = I$ области поиска $D$ (в частном случае, в одной или нескольких ее точках) целевая функция может быть не определена.
+где $L$ -- константа Липщица.}
+\item{В некоторой подобласти $I$ области поиска $D$ (в частном случае, в одной или нескольких ее точках) целевая функция может быть не определена.
 Тогда функция $\phi(y)$ определена и вычислима лишь в некоторой области $Q = D \backslash I$ положительного объёма.
@@
 На случай решения задач с невсюду вычислимой целевой функцией нами была разработана модификация алгоритма глобального поиска (АГП-Н) [\ref{rfa:rulit:Usova2024}].
-Далее будет приведено описание одной из версий модификации алгоритма и указаны условия сходимости данного метода.
+
+\section{Редукция размерности}
+
+Используя кривые типа развертки Пеано, однозначно отображающие отрезок $[0,1]$ на $N$-мерный единичный гиперкуб
+\begin{equation}\label{eq2_} 
+D=\left\{ y \in R^N: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\},
+\end{equation}
+исходную задачу (\ref{eq1}) можно редуцировать к одномерной задаче
+\begin{equation}\label{eq2} 
+f^*(x)=\phi(y(x^* ))=\min_{x \in [0,1]} \left\{ \phi(y(x)) \right\},
+\end{equation}
+что позволяет применить для ее решения эффективные алгоритмы одномерной оптимизации.
+
+Известно, что схема редукции размерности с использованием кривых Пеано сопоставляет многомерной задаче с липшицевой целевой функцией (\ref{eq1}) задачу (\ref{eq2}) с одномерной целевой функцией, удовлетворяющей условию Гёльдера
+\begin{equation}\label{eq4} 
+| f(x')-f(x'') | \leq K \rho(x',x''), \; x',x'' \in [0,1],
+\end{equation}
+где $\rho(x',x'') =  |x' - x''|^{1/N}$ -- метрика Гёльдера, $N$ -- размерность исходной многомерной задачи, а коэффициент $K$ связан с константой Липшица $L$ соотношением $K \leq 2L\sqrt {N+3}$ [\ref{rfa:rulit:Strongin2000}].
 
@@
 Расчёт характеристики интервала с невычислимой и вычислимой точками производить по правилам работы с граничным интервалом.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{pic/fig2.png}
-\caption{Примеры распределения поисковых испытаний в областях невычислимости} \label{fig2}
-\end{figure}
 
@@
 \]
-Если максимальную характеристику имеют несколько интервалов, отдавать предпочтение вычислимым интервалам с минимальным номером.
+Если максимальную характеристику имеют несколько интервалов, выбрать интервал с большим количеством вычислимых точек с минимальным номером.
 
-\textit{Правило 6.} Провести очередное испытание в серединной точке интервала $(x_{t-1},x_t)$, если индексы его концевых точек не совпадают или оба соответсвуют невычислимыми, т.е.
+\textit{Правило 6.} Провести очередное испытание в серединной точке интервала $(x_{t-1},x_t)$, если индексы его концевых точек не совпадают или оба соответсвуют невычислимым точкам, т.е.
 \begin{equation}\label{eq10} 
@@
 
-\textbf{Лемма (о попадании испытаний в область вычислимости)}.
+\textbf{Лемма (о попадании испытаний в вычислимую область)}.
 В последовательности испытаний, порождаемом алгоритмом АГП-Н при минимизации функции $\phi(y), y \in D$, всегда найдется такой номер итерации $h$, что в данной точке целевая функция будет определена и вычислима, т.е.
@@
 
-Установим условия сходимости для АГП-Н и покажем, что расширение алгоритма вычисления характеристики не рождает предельных точек другого типа по сравнению с предельными точками классического АГП, кроме
-%предельных точек областей невычислимости.
+Установим условия сходимости для АГП-Н и покажем, что расширение алгоритма вычисления характеристики не рождает предельных точек другого типа по сравнению с предельными точками классического АГП.
+%, кроме предельных точек областей невычислимости.
 
@@
 Значит наше предположение неверно и либо предельная точка $\overline{y} = y(\overline{x})$ является точкой глобального минимума, либо функция в ней является невычислимой и алгоритмом порождается последовательность невычислимых стягивающихся интервалов с положительной характеристикой, вычисляемой по правилу $R(t) =(1-\frac{1}{r})^2\Delta_t$.
+
+\section{Эксперименты}
+
+\subsection{Примеры решения тестовых задач}
+
+...
+
+\begin{figure}[h]
+\includegraphics[width=450pt]{pic/fig_2.jpg}
+\caption{Распределение точек испытаний при решении задач с различными невычислимыми областями} \label{fig_2}
+\end{figure}
+
+\subsection{Сравнение с другими алгоритмами}
+
+...
 

### 2025-02-04 2025/NumMet/NMP_authorGuidelinesRu_2023.tex
Переработаны разделы Введение, Редукция размерности

@@
 \section{Введение}
-Глобальная оптимизация занимается разработкой методов, призванных решать задачи отыскания точек абсолютного (глобального) минимума многоэкстремальных многомерных функций.
-Возможность достоверной оценки глобального оптимума в таких задачах принципиально основана на наличии априорной информации о решаемой задаче, позволяющей связать возможные значения целевой функции с известными значениями в точках проведенных поисковых испытаний.
 
-Весьма часто такая информация представляется в виде предположения, что целевая функция $\phi(y)$ удовлетворяет условию Липшица с неизвестной априори константой $L$.
-Это предположение можно интерпретировать (применительно к прикладным задачам) как отражение ограниченности мощностей, порождающих изменения в моделируемой системе.
-При этом целевая функция зачастую задаётся в виде ``\textit{черного ящика}, является недифференцируемой и каждое вычисление ее значения в некоторой точке допустимой области может требовать значительных вычислительных ресурсов.
+Процесс выбора оптимальных значений параметров в различных приложениях часто подразумевает решение задачи \textit{глобальной (многоэкстремальной) оптимизации}.
+В таких задачах для целевой функции обычно не известно аналитическое выражение; ее значения вычисляются посредством численного моделирования оптимизируемого объекта (т.е.
+функция задана как «черный ящик»).
+При этом даже одно вычисление значения функции в некоторой точке области поиска (данную процедуру будем называть \textit{поисковым испытанием}) может потребовать существенных ресурсов.
+Типичными примерами задач глобальной оптимизации с функциями вида <<черный ящик>> являются задачи идентификации параметров математических моделей по экспериментальным данным [\colorbox[rgb]{1,1,0}{ссылки}].
 
-Для решения задач липшицевой глобальной оптимизации разработан ряд эффективных детерминированных методов 
+%Широко распространенными математическими моделями процесса выбора оптимальных значений параметров в различных приложениях являются задачи глобальной оптимизации.
+В таких задачах целевая функция обычно не задана аналитическим выражением, ее значения вычисляются программно (т.е.
+функция задана как «черный ящик»).
+При этом даже одно вычисление значения функции в некоторой точке области поиска (в дальнейшем этот процесс будем называть \textit{поисковым испытанием}) может потребовать существенного времени.Типичными примерами задач глобальной оптимизации с функциями вида «черный ящик» являются задачи идентификации параметров математических моделей по экспериментальным данным [ссылки].
+
+Задачи глобальной оптимизации относятся к числу наиболее сложных оптимизационных задач, т.к.
+отыскание абсолютного экстремума требует вычисления значений функции в узлах некоторой адаптивной сетки в области поиска.
+Построение достоверных оценок глобального оптимума в многоэкстремальных задачах принципиально основано на использовании априорной информации, позволяющей связать возможные значения целевой функции с известными значениями в точках проведенных поисковых испытаний.
+Весьма часто такая информация представляется в виде предположения, что целевая функция $\phi(y)$ удовлетворяет в области поиска условию Липшица, причем константа Липшица $L$, характеризующая скорость изменения функции, является неизвестной.
+В прикладных задачах это предположение соответствует условию ограниченности изменения характеристик оптимизируемого объекта при ограниченном изменении его параметров.
+
+Для решения задач липшицевой глобальной оптимизации предложен ряд эффективных алгоритмов 
 [\ref{rfa:rulit:Grishagin2016_2},\ref{rfa:rulit:Jones2021},\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Birect2020},\ref{rfa:rulit:Sergeyev2017}].
-Проведенные сравнения показывают, что детерминированные алгоритмы превосходят (по разным критериям) широко распространенные природно-естественные алгоритмы [\ref{rfa:rulit:Liberti2005},\ref{rfa:rulit:Sergeyev2018}].
+Указанные методы относятся к классу детерминированных и превосходят (в терминах количества испытаний, требующегося для корректного решения задачи) широко распространенные метаэвристические алгоритмы (см.
+результаты сравнения в [\ref{rfa:rulit:Liberti2005},\ref{rfa:rulit:Sergeyev2018}]).
+Данная работа продолжает развитие одного из известных детерминированных алгоритмов решения задач липшицевой глобальной оптимизации -- алгоритма глобального поиска (АГП) [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}].
+В рамках рассматриваемого подхода решение многомерных задач сводится к решению эквивалентных им задач одномерной оптимизации.
+Соответствующая редукция основана на использовании кривых Пеано-Гильберта, однозначно отображающих единичный отрезок вещественной оси на многомерную область поиска.
 
-Данная работа продолжает развитие одного из эффективных детерминированных методов решения задач липшицевой глобальной оптимизации – информационно-статистического алгоритма глобального поиска [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}].
-Данный алгоритм предполагает решение многомерных задач посредством их сведения к задачам одномерной оптимизации при помощи кривой заполняющей пространство (кривые Пеано).
-
-Отметим, что для оптимизации сложных объектов реального мира естественным является использование при расчетах сложных математических моделей, что, как следствие, существенно увеличивает трудоемкость поиска оптимума.
+%Отметим, что для оптимизации сложных объектов реального мира естественным является использование при расчетах сложных математических моделей, что, как следствие, существенно увеличивает трудоемкость поиска оптимума.
 За последние десятилетия специалистами в области оптимизации и параллельных вычислений было предложено множество способов снижения вычислительной сложности и ускорения используемых алгоритмов, связанных как с решением возникающих оптимизационных задач [\ref{rfa:rulit:Kvasov2013},\ref{rfa:rulit:Sergeyev2020}], так и с численным анализом исходных моделей [\ref{rfa:rulit:Dongarra2022},\ref{rfa:rulit:Duwe2020}].
-
-Однако в последнее время актуальной становится принципиально новая проблема прикладных задач – численная нестабильность исследуемых моделей в некоторых (заранее не известных) подобластях области изменения параметров.
-В таких подобластях невозможно корректно провести численное моделирование и вычислить значение целевой функции задачи.
-Данное явление можно интерпретировать или как наличие в задаче некоторых скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], или как наличие неизвестных разрывных областей [\ref{rfa:rulit:Audet2022}], или как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Sergeyev2003},\ref{rfa:rulit:Strongin2020}].
+В последнее время при решении прикладных оптимизационных задач актуальной становится проблема численной нестабильности исследуемых моделей в некоторых (заранее не известных) подобластях области изменения параметров [\colorbox[rgb]{1,1,0}{Модорский?}].
+В таких подобластях невозможно корректно провести численное моделирование и вычислить значение оптимизируемого критерия.
+Данное свойство можно интерпретировать как наличие в задаче скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], либо как существование  заранее неизвестных областей неопределенности целевой функции [\ref{rfa:rulit:Audet2022}], либо как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Sergeyev2003},\ref{rfa:rulit:Strongin2020}].
 В такой постановке задача оптимизации существенно усложняется, т.к.
 область допустимых сочетаний параметров является заранее неопределенной.
+Новое предложение, исследуемое в данной статье, связано с модификацией правил алгоритма глобального поиска для корректного решения задач, в которых целевая функция может быть определена не во всей области поиска.
 
-В статье приводится доказательство сходимости одного из предложенных ранее подходов к модификации алгоритма глобального поиска на случай частично определенной целевой функции [\ref{rfa:rulit:Usova2024}].
-Также продемонстрированы новые результаты численных экспериментов.
+Статья имеет следующую структуру.
+Первый раздел содержит постановку задачи и описание используемого способа редукции размерности.
+Вычислительные правила модифицированного алгоритма глобального поиска, обеспечивающего работу с частично определенными целевыми функциями, приведены во втором разделе.
+В третьем разделе доказаны теоретические утверждения о сходимости предложенного алгоритма.
+В четвертом разделе приведено описание генератора, порождающего тестовые задачи с не всюду определенной целевой функцией.
+Использование генератора задач позволяет оценить эффективность предложенного алгоритма путем численного решения большой серии задач, в том числе -- провести его сравнение с известными методами глобальной оптимизации.
+В заключительном разделе обсуждаются полученные результаты и формулируются направления дальнейших исследований.
 
@@
 \end{equation}
-где $y=(y_1,y_2,...,y_N)$ -– вектор варьируемых параметров, $D$ -- $N$-мерный гиперкуб, $N$ –- размерность решаемой задачи.
+где $y=(y_1,y_2,...,y_N)$ -- вектор варьируемых параметров, $D$ -- $N$-мерный гиперкуб, $N$ –- размерность решаемой задачи.
 О целевой функции $\phi (y)$ мы делаем следующие предположения.
@@
 
-Последнее предположение делает невозможными применение классического информационно-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска  [\ref{rfa:rulit:Strongin2000}] или других методов липшицевой оптимизации [\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Sergeyev2017}].
-На случай решения задач с невсюду вычислимой целевой функцией нами была разработана модификация алгоритма глобального поиска (АГП-Н) [\ref{rfa:rulit:Usova2024}].
+Последнее предположение делает невозможными применение классического информационно-статистического алгоритма глобального поиска  [\ref{rfa:rulit:Strongin2000}] или других методов липшицевой оптимизации [\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Sergeyev2017}].
+%На случай решения задач с невсюду вычислимой целевой функцией нами была разработана модификация алгоритма глобального поиска (АГП-Н) [\ref{rfa:rulit:Usova2024}].
 
@@
 
-Используя кривые типа развертки Пеано, однозначно отображающие отрезок $[0,1]$ на $N$-мерный единичный гиперкуб
+Многие методы липшицевой глобальной оптимизации для решения многомерных задач используют идею редукции размерности.
+Указанная редукция может быть основана, например, на построении диагональных кривых [\colorbox[rgb]{1,1,0.6}{Квасов}], схеме вложенной оптимизации [\colorbox[rgb]{1,1,0.6}{Гришагин}], симплексном разбиении области поиска [\colorbox[rgb]{1,1,0.6}{Жилинскас}], и т.п.
+В нашей работе для редукции размерности используются кривые Пеано-Гильберта $y(x)$, однозначно отображающие отрезок вещественной оси  $[0,1]$ на $N$-мерный единичный гиперкуб
 \begin{equation}\label{eq2_} 
-D=\left\{ y \in R^N: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\},
+D=\left\{ y \in R^N: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\}.
 \end{equation}
-исходную задачу (\ref{eq1}) можно редуцировать к одномерной задаче
+С их помощью исходную задачу (\ref{eq1}) можно свести к задаче 
 \begin{equation}\label{eq2} 
-f^*(x)=\phi(y(x^* ))=\min_{x \in [0,1]} \left\{ \phi(y(x)) \right\},
+f^*(x)=\phi(y(x^* ))=\min_{x \in [0,1]} \left\{ \phi(y(x)) \right\}
 \end{equation}
-что позволяет применить для ее решения эффективные алгоритмы одномерной оптимизации.
+и применить для ее решения эффективные алгоритмы одномерной оптимизации.
 
-Известно, что схема редукции размерности с использованием кривых Пеано сопоставляет многомерной задаче с липшицевой целевой функцией (\ref{eq1}) задачу (\ref{eq2}) с одномерной целевой функцией, удовлетворяющей условию Гёльдера
+Отметим, что кривая Пеано-Гильберта определяется как предельный объект, поэтому на практике можно использовать лишь ее аппроксимацию (\textit{развертку}) $y_m(x)$, приближающую истинную кривую $y(x)$ с точностью $2^{-m}$ (целое число $m \geq 1$ является параметром построения развертки).
+Различные способы численного построения разверток рассмотрены в [\ref{rfa:rulit:Strongin2000}].
+
+Также известно, что схема редукции размерности с использованием разверток сопоставляет многомерной липшицевой функции $\phi(y)$ одномерную функцию $f(x)$, удовлетворяющую условию Гёльдера
 \begin{equation}\label{eq4} 
-| f(x')-f(x'') | \leq K \rho(x',x''), \; x',x'' \in [0,1],
+| f(x')-f(x'') | \leq H \rho(x',x''), \; x',x'' \in [0,1],
 \end{equation}
-где $\rho(x',x'') =  |x' - x''|^{1/N}$ -- метрика Гёльдера, $N$ -- размерность исходной многомерной задачи, а коэффициент $K$ связан с константой Липшица $L$ соотношением $K \leq 2L\sqrt {N+3}$ [\ref{rfa:rulit:Strongin2000}].
+где $\rho(x',x'') =  |x' - x''|^{1/N}$ -- метрика Гёльдера, $N$ -- размерность исходной многомерной задачи, а коэффициент $H$ связан с константой Липшица $L$ соотношением $H \leq 2L\sqrt {N+3}$ [\ref{rfa:rulit:Strongin2000}].
+
+Таким образом, в редуцированной одномерной задаче (\ref{eq2}) поисковое испытание в точке $x'$ подразумевает сначала построение образа $y' = y(x')$, и лишь затем -- вычисление значения $z'=\phi(y')$; результатом испытания будет являться пара значений $(x',z')$.
 

### 2025-02-04 2025/NumMet/NMP_authorGuidelinesRu_2023.tex
переработан раздел Постановка задачи

@@
 \section{Постановка задачи}
+
 В общем виде задача глобальной оптимизации может быть сформулирована следующим образом:
@@
 где $y=(y_1,y_2,...,y_N)$ -- вектор варьируемых параметров, $D$ -- $N$-мерный гиперкуб, $N$ –- размерность решаемой задачи.
-О целевой функции $\phi (y)$ мы делаем следующие предположения.
 
+Будем предполагать, что целевая функция $\phi (y)$:
 \begin{enumerate}
-\item{Целевая функция может быть многоэкстремальной, недифференцируемой и, более того, заданной в форме ``черного ящика'' (т.е.
-в виде некоторой подпрограммы, на вход которой подается аргумент, а выходом является соответствующее значение функции).}
-\item{Каждое вычисление функции в некоторой точке допустимой области может требовать значительных вычислительных ресурсов.}
-\item{Целевая функция удовлетворяет условию Липщица
-\begin{equation}\label{eq3} 
-| \phi (y')-\phi (y'') | \leq L \| y'-y'' \|, \; y',y'' \in D,
-\end{equation}
-где $L$ -- константа Липщица.}
-\item{В некоторой подобласти $I$ области поиска $D$ (в частном случае, в одной или нескольких ее точках) целевая функция может быть не определена.
-Тогда функция $\phi(y)$ определена и вычислима лишь в некоторой области $Q = D \backslash I$ положительного объёма.
-Отметим, что исходя из опыта решения прикладных оптимизационных задач [\ref{rfa:rulit:Barkalov2022},\ref{rfa:rulit:Gubaydullin2022}], суммарный объем области невычислимости $I$ составляет небольшую долю объема области поиска $D$.}
+\item{является многоэкстремальной и, возможно, недифференцируемой;}
+\item {задана как <<черный ящик>> -- в виде некоторой вычислительной процедуры, на вход которой подается аргумент, а выходом является соответствующее значение функции;}
+\item{является вычислительно трудоемкой, т.е. проведение одного поискового испытания может потребовать существенного времени;}
+\item{определена и вычислима только в некоторой заранее неизвестной подобласти $Q \subseteq D$ положительного объема.}
 \end{enumerate}
 
-Последнее предположение делает невозможными применение классического информационно-статистического алгоритма глобального поиска  [\ref{rfa:rulit:Strongin2000}] или других методов липшицевой оптимизации [\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Sergeyev2017}].
+%методов липшицевой оптимизации [\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Sergeyev2017}].
 %На случай решения задач с невсюду вычислимой целевой функцией нами была разработана модификация алгоритма глобального поиска (АГП-Н) [\ref{rfa:rulit:Usova2024}].
 
-\section{Редукция размерности}
+Подобласть $Q \subseteq D$ будем называть \textit{областью вычислимости} функции $\phi (y)$. 
+Соответственно, подобласть $I = D \backslash Q$ будем называть \textit{областью невычислимости} функции $\phi (y)$.
+Отметим, что исходя из опыта решения прикладных оптимизационных задач [\ref{rfa:rulit:Barkalov2022},\ref{rfa:rulit:Gubaydullin2022}], суммарный объем области невычислимости $I$ составляет небольшую долю объема области поиска $D$.
+
+Предположение 4 делает невозможными использование базового АГП [\ref{rfa:rulit:Strongin2000}] или его параллельных вариантов [\colorbox[rgb]{1,1,0.6}{ссылки}] для решения задачи (\ref{eq1}).
+Чтобы преодолеть это ограничение, будем предполагать, что существует функция $\Phi (y)$, вычислимая во всей области $D$ и удовлетворяющая условию Липшица
+\begin{equation}\label{eq3} 
+| \Phi (y')-\Phi (y'') | \leq L \| y'-y'' \|, \; y',y'' \in D,
+\end{equation}
+причем $\phi (y) = \Phi (y), y \in Q$. Функцию $\Phi (y)$ будем называть \textit{липшицевым продолжением} $\phi (y)$ на область $D$. 
+Существование функции $\Phi (y)$ позволяет разработать модифицированный алгоритм глобального поиска для решения задач с частично вычислимыми функциями, детальное описание которого приведено в разделе \ref{sec:algorithm}.
+
+\section{Редукция размерности}\label{sec:reduction}
 
 Многие методы липшицевой глобальной оптимизации для решения многомерных задач используют идею редукции размерности.
-Указанная редукция может быть основана, например, на построении диагональных кривых [\colorbox[rgb]{1,1,0.6}{Квасов}], схеме вложенной оптимизации [\colorbox[rgb]{1,1,0.6}{Гришагин}], симплексном разбиении области поиска [\colorbox[rgb]{1,1,0.6}{Жилинскас}], и т.п.
+Указанная редукция может быть основана, например, на построении диагональных кривых [\colorbox[rgb]{1,1,0.6}{Квасов}], схеме вложенной оптимизации [\colorbox[rgb]{1,1,0.6}{Гришагин}], симплексном разбиении области поиска [\colorbox[rgb]{1,1,0.6}{Жилинскас}].
 В нашей работе для редукции размерности используются кривые Пеано-Гильберта $y(x)$, однозначно отображающие отрезок вещественной оси  $[0,1]$ на $N$-мерный единичный гиперкуб
@@
 
-\section{Алгоритм решения задач с частично определенной целевой функцией}
+\section{Алгоритм решения задач с частично определенной целевой функцией}\label{sec:algorithm}
 

### 2025-02-04 2025/NumMet/NMP_authorGuidelinesRu_2023.tex
fix the list of papers

@@
 Указанные методы относятся к классу детерминированных и превосходят (в терминах количества испытаний, требующегося для корректного решения задачи) широко распространенные метаэвристические алгоритмы (см.
-результаты сравнения в [\ref{rfa:rulit:Liberti2005},\ref{rfa:rulit:Sergeyev2018}]).
+результаты сравнения в [\ref{rfa:rulit:Sovrasov2019},\ref{rfa:rulit:Sergeyev2018}]).
 Данная работа продолжает развитие одного из известных детерминированных алгоритмов решения задач липшицевой глобальной оптимизации -- алгоритма глобального поиска (АГП) [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}].
@@
 }
-%14
+%14  - замена на русскую
  \rlitem{rfa:rulit:Sergeyev2017}{%
-\textit{Sergeyev~Y.D., Kvasov~D.E.}
-Deterministic global optimization: An introduction to the diagonal approach.
-New York: Springer, 2017.
-\doi{10.1007/978-1-4939-7199-2}.
+\textit{Сергеев~Я.Д., Квасов~Д.Е.}
+Диагональные методы глобальной оптимизации.
+М.: Физматлит, 2008.
 }
-%11
-\rlitem{rfa:rulit:Liberti2005}{%
-\textit{Liberti~L., Kucherenko~S.}
-Comparison of deterministic and stochastic approaches to global optimization~// 
-Int. Trans. Oper. Res. 2005. \textbf{12}. 263--285.
+%11 - замена
+\rlitem{rfa:rulit:Sovrasov2019}{%
+\textit{Sovrasov~V.}
+Comparison of several stochastic and deterministic derivative-free global optimization algorithms~// 
+Lecture notes in computer science. 2019. \textbf{11548}. 70--81.
 }
@@
 }
-%10
+%10 - замена на русскую
 \rlitem{rfa:rulit:Kvasov2013}{%
-\textit{Kvasov~D.E., Sergeyev~Y.D.}
-Lipschitz global optimization methods in control problems~// 
-Autom. Remote Control. 2013. \textbf{74}, N~9. 1435--1448.
-\doi{10.1134/s0005117913090014}.
+\textit{Квасов~Д.Е., Сергеев~Я.Д.}
+Методы липшицевой глобальной оптимизации в задачах управления~// 
+Автоматика и телемеханика. 2013. №~9. 3--19.
 }

### 2025-02-05 2025/NumMet/NMP_authorGuidelinesRu_2023.tex
Исправлен раздел с описанием алгоритма

@@
 За последние десятилетия специалистами в области оптимизации и параллельных вычислений было предложено множество способов снижения вычислительной сложности и ускорения используемых алгоритмов, связанных как с решением возникающих оптимизационных задач [\ref{rfa:rulit:Kvasov2013},\ref{rfa:rulit:Sergeyev2020}], так и с численным анализом исходных моделей [\ref{rfa:rulit:Dongarra2022},\ref{rfa:rulit:Duwe2020}].
-В последнее время при решении прикладных оптимизационных задач актуальной становится проблема численной нестабильности исследуемых моделей в некоторых (заранее не известных) подобластях области изменения параметров [\colorbox[rgb]{1,1,0}{Модорский?}].
+В последнее время при решении прикладных оптимизационных задач актуальной становится проблема численной нестабильности исследуемых моделей в некоторых (заранее не известных) подобластях области изменения параметров [\colorbox[rgb]{1,1,0}{ссылка}].
 В таких подобластях невозможно корректно провести численное моделирование и вычислить значение оптимизируемого критерия.
-Данное свойство можно интерпретировать как наличие в задаче скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], либо как существование  заранее неизвестных областей неопределенности целевой функции [\ref{rfa:rulit:Audet2022}], либо как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Sergeyev2003},\ref{rfa:rulit:Strongin2020}].
+Данное свойство можно интерпретировать как наличие в задаче скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], либо как существование  заранее неизвестных областей неопределенности целевой функции [\ref{rfa:rulit:Audet2022}], либо как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Strongin2020}].
 В такой постановке задача оптимизации существенно усложняется, т.к.
@@
 Многие методы липшицевой глобальной оптимизации для решения многомерных задач используют идею редукции размерности.
-Указанная редукция может быть основана, например, на построении диагональных кривых [\colorbox[rgb]{1,1,0.6}{Квасов}], схеме вложенной оптимизации [\colorbox[rgb]{1,1,0.6}{Гришагин}], симплексном разбиении области поиска [\colorbox[rgb]{1,1,0.6}{Жилинскас}].
+Указанная редукция может быть основана, например, на построении диагональных кривых [\ref{rfa:rulit:Sergeyev2017}], схеме вложенной оптимизации [\ref{rfa:rulit:Grishagin2016_2}], симплексном разбиении области поиска [\ref{rfa:rulit:PaulaviciusZilinskas2014}].
 В нашей работе для редукции размерности используются кривые Пеано-Гильберта $y(x)$, однозначно отображающие отрезок вещественной оси  $[0,1]$ на $N$-мерный единичный гиперкуб
@@
 
-Таким образом, в редуцированной одномерной задаче (\ref{eq2}) поисковое испытание в точке $x'$ подразумевает сначала построение образа $y' = y(x')$, и лишь затем -- вычисление значения $z'=\phi(y')$; результатом испытания будет являться пара значений $(x',z')$.
-
 \section{Алгоритм решения задач с частично определенной целевой функцией}\label{sec:algorithm}
 
-Для реализации модификации алгоритма рассматрвается простейший подход - игнорировать отсутствие значения целевой функции в точке.
-Расчёт характеристики интервала с двумя невычислимыми или с невычислимой и граничной точками производить по длине интервала с использованием понижающей константы.
-Расчёт характеристики интервала с невычислимой и вычислимой точками производить по правилам работы с граничным интервалом.
+Предлагаемый алгоритм представляет собой модификацию алгоритма глобального поиска (АГП) и ориентирован на решение задач, в которых целевая функция определена и вычислима только в области $Q \subseteq D$, причем область $Q$ является априори неизвестной.
+В случае всюду вычислимой целевой функции, т.к.
+когда $Q=D$, результаты работы нового алгоритма должны соответствовать исходному АГП.
+Изменения вычислительных правил алгоритма касаются только учета и обработки результатов испытаний, полученных в точках из области невычислимости $I$.
 
-Приведем подробное описание модификации алгоритма глобального поиска для невсюду вычислимой целевой функции (АГП-Н) для решения задач (\ref{eq2}).
+Отметим, что в редуцированной одномерной задаче (\ref{eq2}) поисковое испытание в точке $x' \in [0,1]$ предполагает сначала построение образа $y' = y(x') \in D$, и лишь затем -- вычисление значения $z'=\phi(y')$.
+Результатом испытания будет являться набор $(x',y',z')$, причем значение $z'$ будет являться неопределенным, если $y' \in I$.
+Для исключения обработки неопределенных значений в алгоритме вводится понятие \textit{индекса} точки испытания (см.
+Правило 1 ниже).
 
-\subsection{Описание алгоритма АГП-Н}
+Приведем подробное описание нового алгоритма, на который будем ссылаться как на Алгоритм Глобального Поиска для задач с Не всюду вычислимой целевой функцией, АГП-Н.
 
-На каждой итерации глобального поиска выполняется \textit{испытание}.
-Испытанием будем называть вычисление значения оптимизируемой функции $\phi (y(x))$ из (\ref{eq2}).
-Первое испытание будет осуществляться в серединной внутренней точке $x^1 \in (0,1)$.
-Выбор точки $x^{k+1}, k \geq 1$ очередного $(k+1)^\text{-го}$ испытания осуществляется на основе следующих правил.
+\subsection{Вычислительная схема алгоритма}
+
+Первые два испытания проводятся в граничных точках $x^0=0$ и $x^0=1$.
+Выбор точки $x^{k+1}, k \geq 1$ очередного $(k+1)$-го испытания осуществляется в соответствии со следующими правилами.
 
@@
 \end{equation}
-и сопоставить им значения $z_i=f(x_i), 0 < i < k$, вычисленные в этих точках, и индекс $v_i=v(x_i)$, определяемый по правилу
+и сопоставить каждой точке значение $z_i=f(x_i)$ и индекс $v_i=v(x_i)$, определяемый по правилу
 \begin{equation}\label{eq6} 
@@
   \begin{cases}
-    -1, & {\quad \text{if } x_i \text{ -- граничная точка}},\\
-    0, & {\quad \text{if } x_i \text{ -- невычислимая точка}},\\
-    1, & {\quad \text{if } x_i \text{ -- внутренняя точка}}.
+    0, & {\quad y(x_i) \in I },\\
+    1, & {\quad y(x_i) \in Q}.
   \end{cases}
 \end{equation}
-Точки $x_0=0$ и $x_{k}=1$ введены дополнительно (значения $z_0$ и $z_{k}$ не определены).
+%ссылка на множества Q и I.
 
-\textit{Правило 2.} Вычислить текущую нижнюю оценку
+\textit{Правило 2.} Вычислить величину
 \begin{equation}\label{eq7} 
-\mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i},\quad v(x_{i-1})=v(x_i)=1, \quad 2 \leq i \leq k  \right\},\quad \Delta _i= (x_i-x_{i-1})^{1/N},
+\mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i},\quad v(x_{i-1})=v(x_i)=1, \quad 1 \leq i \leq k  \right\},\quad \Delta _i= (x_i-x_{i-1})^{1/N}.
 \end{equation}
-для константы Гёльдера $K$ из (\ref{eq4}) для редуцированной функции $f(x)$, где $r>1$ -- параметр надёжности. Если $\mu=0$, то принять $\mu=1$.
+Если $\mu=0$, то положить $\mu=1$.
 
-\textit{Правило 3.} Определить текущее лучшее значение целевой функции
+\textit{Правило 3.} Определить текущее минимальное значение целевой функции
 \begin{equation}\label{eq8} 
-z^*= \min \left\{ f(x_i): v(x_i)=1, 0 < i < k \right\}.
+z^*= \min \left\{ z_i: v(x_i)=1, 0 \leq i \leq k \right\}.
 \end{equation}
@@
   \begin{cases}
-    {(1-\frac{1}{r})}^2 \Delta _i,& {\quad v(x_i)=v(x_{i-1})=0 \text{ или } v(x_i)=0 \text{ и } v(x_{i-1})=-1}\\
-    {} & {\quad \text{или } v(x_i)=-1  \text{ и } v(x_{i-1})=0},\\
+    {(1-\frac{1}{r})}^2 \Delta _i,& {\quad v(x_i)=v(x_{i-1})=0} \\
     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1},\\
-    2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad  v(x_{i-1})=-1 \text{ или } v(x_{i-1})=0, v(x_i)=1},\\
-    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1 \text{ или } v(x_i)=0.}
+    2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad v(x_{i-1})=0, v(x_i)=1},\\
+    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0.}
   \end{cases}
@@
 
-\textit{Правило 5.} Выбрать интервал $(x_{t-1},x_t)$ с максимальным значением характеристики $R(i)$:
+\textit{Правило 5.} Определить интервал $(x_{t-1},x_t)$ с максимальным значением характеристики:
 \[
@@
 \]
-Если максимальную характеристику имеют несколько интервалов, выбрать интервал с большим количеством вычислимых точек с минимальным номером.
+Если максимальную характеристику имеют несколько интервалов, выбрать интервал
+%с большим количеством вычислимых точек - это можно формализовать как интеврал с наибольшей суммой индексов граничных точек
+с минимальным номером.
 
-\textit{Правило 6.} Провести очередное испытание в серединной точке интервала $(x_{t-1},x_t)$, если индексы его концевых точек не совпадают или оба соответсвуют невычислимым точкам, т.е.
-\begin{equation}\label{eq10} 
-x^{k+1}=\frac {x_t+x_{t-1}}{2},\quad v(x_{t-1})\neq v(x_t) \text{ или } v(x_{t-1})=v(x_t)=0.
-\end{equation}
-В противном случае провести испытание в точке 
+\textit{Правило 6.} Если $v(x_i)=v(x_{i-1})=1$, то провести очередное испытание в точке 
 \begin{equation}\label{eq11} 
@@
 \end{equation}
+В противном случае провести очередное испытание в серединной точке интервала $(x_{t-1},x_t)$, т.е.
+\begin{equation}\label{eq10} 
+x^{k+1}=\frac {x_t+x_{t-1}}{2}.
+\end{equation}
 
-\textit{Условие остановки.} Поиск завершен, если длина интервала $(x_{t-1},x_t)$, содержащего очередную точку $x^{k+1}$, не превышает заданной точности $\varepsilon$, то есть $\Delta _t \leq \varepsilon$, где $t$ из Правила 5, а $\varepsilon>0$ -- задаваемый параметр алгоритма.
+\textit{Условие остановки.} Поиск завершается при выполнении условия 
+\[
+\Delta _t = (x_t-x_{t-1})^{1/N} \leq \varepsilon,
+\]
+где $t$ из Правила 5, а $\varepsilon>0$ -- параметр алгоритма (точность поиска).
 
@@
 
-%7
+%7 - замена
 \rlitem{rfa:rulit:Grishagin2016_2}{%
-\textit{Grishagin~V.A., Israfilov~R.A.}
-Global search acceleration in the nested optimization scheme~// 
-AIP Conference Proceedings. 2016. \textbf{1738}. 400010.
-\doi{10.1063/1.4952198}.
+\textit{Grishagin~V.A., Israfilov~R.A., Sergeyev~Y.D.}
+Convergence conditions and numerical comparison of global optimization methods based on dimensionality reduction schemes~// 
+Appl. Math. Comput. 2018. \textbf{318}. 270--280.
+\doi{10.1016/j.amc.2017.06.036}.
 }
@@
 }
-%4
+%4 - убрать
 \rlitem{rfa:rulit:Dongarra2022}{%
@@
 }
-%5
+%5 - убрать
 \rlitem{rfa:rulit:Duwe2020}{%
@@
 }
-%18
+%18 - убрать
 \rlitem{rfa:rulit:Sergeyev2003}{%

### 2025-02-06 2025/NumMet/NMP_authorGuidelinesRu_2023.tex
исправлена первая лемма

@@
 За последние десятилетия специалистами в области оптимизации и параллельных вычислений было предложено множество способов снижения вычислительной сложности и ускорения используемых алгоритмов, связанных как с решением возникающих оптимизационных задач [\ref{rfa:rulit:Kvasov2013},\ref{rfa:rulit:Sergeyev2020}], так и с численным анализом исходных моделей [\ref{rfa:rulit:Dongarra2022},\ref{rfa:rulit:Duwe2020}].
-В последнее время при решении прикладных оптимизационных задач актуальной становится проблема численной нестабильности исследуемых моделей в некоторых (заранее не известных) подобластях области изменения параметров [\colorbox[rgb]{1,1,0}{ссылка}].
-В таких подобластях невозможно корректно провести численное моделирование и вычислить значение оптимизируемого критерия.
-Данное свойство можно интерпретировать как наличие в задаче скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], либо как существование  заранее неизвестных областей неопределенности целевой функции [\ref{rfa:rulit:Audet2022}], либо как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Strongin2020}].
+В последнее время при решении прикладных оптимизационных задач актуальной становится проблема численной нестабильности исследуемых моделей в некоторых (заранее не известных) подобластях области изменения параметров.
+В таких подобластях невозможно корректно провести численное моделирование и вычислить значение целевого критерия.
+Например, подобные эффекты возникают при численном моделировании обледенения лопаток вентилятора газотурбинного двигателя [\ref{rfa:rulit:Modorskii}].
+Данное свойство можно интерпретировать как наличие в оптимизационной задаче скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], либо как существование  заранее неизвестных областей неопределенности целевой функции [\ref{rfa:rulit:Audet2022}], либо как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Strongin2020}].
 В такой постановке задача оптимизации существенно усложняется, т.к.
@@
   \begin{cases}
-    {(1-\frac{1}{r})}^2 \Delta _i,& {\quad v(x_i)=v(x_{i-1})=0} \\
+    {(1-\frac{1}{r})}^2 \Delta _i,& {\quad v(x_i)=v(x_{i-1})=0,} \\
     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1},\\
     2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad v(x_{i-1})=0, v(x_i)=1},\\
-    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0.}
+    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0,}
   \end{cases}
 \end{equation}
-
+где $r>1$ -- параметр метода.
 \textit{Правило 5.} Определить интервал $(x_{t-1},x_t)$ с максимальным значением характеристики:
@@
 
+%А нужна ли здесь формулировка теоремы о сходимости АГП? Можно только сослаться на нее
 \textbf{Теорема (о сходимости АГП)}.
 Пусть $\overline{y}=y(\overline{x})$ есть предельная точка последовательности испытаний $\{y^k\}=\{y(x^k)\}, \{x^k\} \subset [0,1]$, порождаемой АГП при минимизации липшицевой с константой $L$ функции $\phi(y), y = y(x) \in D$.
-Тогда если на некотором шаге поиска выполнено условие
+Тогда:
+
+\begin{enumerate}
+\item
+если у последовательности $\{y^k\}$ одновременно с $\overline{y}$ существует другая предельная точка $y'$, то  $\phi(\overline{y}) = \phi(y')$.
+\item
+для любого $k \geq 0$ выполняется неравенство $z^k = \phi(y^k) \geq \phi(\overline{y})$
+\item
+если на некотором шаге поиска выполнено условие 
 \[
@@
 то любая точка глобального минимума $y^*$ является предельной точкой последовательности $\{y^k\}$ и любая точка $\overline{y}$ является точкой глобального минимума функции $\phi(y)$.
+\end{enumerate}
 
@@
 
-\textbf{Лемма (о попадании испытаний в вычислимую область)}.
-В последовательности испытаний, порождаемом алгоритмом АГП-Н при минимизации функции $\phi(y), y \in D$, всегда найдется такой номер итерации $h$, что в данной точке целевая функция будет определена и вычислима, т.е.
-$v(x^h) = 1$.
+%лемма использует утверждение о том, что R(i)>0, т.е. ее нельзя доказывать раньше теоремы, только внутри нее
+\textbf{Лемма} (о попадании испытаний в область вычислимости).
+Пусть $\{y^k\}=\{y(x^k)\}$ -- последовательность точек испытаний, порождаемая алгоритмом АГП-Н при минимизации частично определенной в области $D$ фукнции $\phi(y)$.
+Тогда существует номер $h \geq 0$ такой, что точка $y^h = y(x^h)$ принадлежит области вычислимости, т.е.
+$y^h \in Q$ и $v(x^h) = 1$.
 
 \textbf{Доказательство}.
-Пусть $y' \in Q$ -- внутренняя точка множества $Q$.
-Значит существует такая точка $x' \in [0,1]$, являющаяся прообразом $y' = y(x')$, и такой отрезок $[\alpha, \beta] \subset [0, 1]$, содержащий точку $x'$, что для любой точки $x \in [\alpha, \beta]$ ее образ попадает в область вычислимости $y(x) \in Q$ (Рис.
+Пусть $y' \in Q$ -- внутренняя точка множества $Q$, а $x' \in [0,1]$ -- ее прообраз на отрезке $[0,1]$, т.е.
+$y' = y(x')$.
+Тогда существует такой отрезок $[\alpha, \beta] \subset [0, 1]$, содержащий точку $x'$, что для любой точки $x \in [\alpha, \beta]$ ее образ $y(x)$ попадает в область вычислимости $Q$, т.е.
+$y(x) \in Q$ (Рис.
 \ref{fig_1}).
@@
 \item{$\phi(y)$ определена и вычислима в области $Q = D \backslash I$ положительного объёма, где $I$ -- область невычислимости;}
-\item{$\phi(y)$ допускает липшицево с константой $L$ продолжение $\overline{\phi}(y)$ на область $D$, т.е.
+\item{$\phi(y)$ допускает липшицево с константой $L$ продолжение $\Phi(y)$ на область $D$, т.е.
 \[
-\phi(y(x))=\overline{\phi}(y(x)), y(x) \in Q.
+\phi(y(x))=\Phi(y(x)), y(x) \in Q.
 \]
@@
 }
+%Новая статья
+\rlitem{rfa:rulit:Modorskii}{%
+\textit{Калюлин~С.Л., Саженков~Н.А., Модорский~В.Я., Владимиров~Н.В.}
+Численное моделирование газодинамических и прочностных характеристик вентилятора для экспериментальной установки по исследованию разрушения льда на вращающихся рабочих лопатках ~// 
+Вестник Пермского национального исследовательского политехнического университета. Механика. 2023. №~1. 134--141.
+\doi{10.14529/jsfi200101}.
+}
+
 %4 - убрать

### 2025-02-14 2025/NumMet/NMP_authorGuidelinesRu_2023.tex
исправлено доказательство теоремы

@@
   \begin{cases}
-    {(1-\frac{1}{r})}^2 \Delta _i,& {\quad v(x_i)=v(x_{i-1})=0,} \\
+    \alpha {(1-\frac{1}{r})}^2 \Delta _i,& {\quad v(x_i)=v(x_{i-1})=0,} \\
     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1},\\
@@
 \end{equation}
-где $r>1$ -- параметр метода.
+где $0 \leq \alpha \leq 1$ и $r>1$ -- параметры метода.
+
 \textit{Правило 5.} Определить интервал $(x_{t-1},x_t)$ с максимальным значением характеристики:
-\[
+\begin{equation}\label{Rt} 
 R(t)= \max\{R(i): \; 1 \leq i \leq k\}.
-\]
+\end{equation}
 Если максимальную характеристику имеют несколько интервалов, выбрать интервал
@@
 Различные варианты алгоритмов для решения задач липшицевой оптимизации и соответствующая теория сходимости представлены в работах [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}]. 
-Приведём формулировку теоремы сходимости для АГП.
+%Приведём формулировку теоремы сходимости для АГП.
 
@@
 
-\textbf{Замечание}. Теорема соответствует случаю пустой области невычислимости целевой функции, т.е. $I = \emptyset$ и $Q = D$.
-
 %лемма использует утверждение о том, что R(i)>0, т.е. ее нельзя доказывать раньше теоремы, только внутри нее
@@
 Пусть $\{y^k\}=\{y(x^k)\}$ -- последовательность точек испытаний, порождаемая алгоритмом АГП-Н при минимизации частично определенной в области $D$ фукнции $\phi(y)$.
-Тогда существует номер $h \geq 0$ такой, что точка $y^h = y(x^h)$ принадлежит области вычислимости, т.е.
+Тогда существует номер $h \geq 1$ такой, что точка $y^h = y(x^h)$ принадлежит области вычислимости, т.е.
 $y^h \in Q$ и $v(x^h) = 1$.
@@
 Тогда существует такой отрезок $[\alpha, \beta] \subset [0, 1]$, содержащий точку $x'$, что для любой точки $x \in [\alpha, \beta]$ ее образ $y(x)$ попадает в область вычислимости $Q$, т.е.
-$y(x) \in Q$ (Рис.
-\ref{fig_1}).
+$y(x) \in Q$.
+Для иллюстрации на Рис.
+\ref{fig_1} \colorbox[rgb]{1,1,0}{ДАТЬ ОПИСАНИЕ РИСУНКА}.
 
@@
 
-Значит на каждом шаге поиска $k \geq 1$ существует некоторый интервал $(x_{j-1}, x_{j})$ такой, что $x' \in (x_{j-1}, x_{j})$.
-Если $y(x_{j-1})$ или $y(x_{j})$ принадлежат области $Q$, то $v(x_{j-1}) = 1$ или $v(x_j) = 1$ и всё доказано.
+Значит, на каждом шаге поиска $k \geq 1$ существует некоторый интервал $(x_{j-1}, x_{j})$ такой, что $x' \in (x_{j-1}, x_{j})$.
+Если $y(x_{j-1})$ или $y(x_{j})$ принадлежат области $Q$, то $v(x_{j-1}) = 1$ или $v(x_j) = 1$, что доказывает утверждение.
 
-Иначе, $R(j)=(1-\frac{1}{r})^2\Delta_j>0$ и в этот интервал попадёт точка одного из следующих испытаний.
-При $k \to \infty$ алгоритм будет порождать последовательность интервалов, сходящихся к $x'$, а значит через некоторое количество итераций $h$ граница одного из интервалов попадёт в $[\alpha, \beta]$ и $v(x_{j+h-1}) = 1$ или $v(x_{j+h}) = 1$, что доказывает утверждение.
+Иначе, в соответствии с (\ref{eq10}), данный интервал будет иметь положительную характеристику $R(j)=\alpha(1-\frac{1}{r})^2\Delta_j>0$ и будет разделел пополам точкой одного из следующих испытаний.
+При $k \to \infty$ алгоритм будет порождать последовательность интервалов, сходящихся к $x'$.
+Следовательно, через некоторое количество итераций $h$ граница одного из интервалов попадёт в $[\alpha, \beta]$,
+%т.е. будут выполняться условия $v(x_{j+h-1}) = 1$ или $v(x_{j+h}) = 1$, 
+что также доказывает утверждение.
 
-Установим условия сходимости для АГП-Н и покажем, что расширение алгоритма вычисления характеристики не рождает предельных точек другого типа по сравнению с предельными точками классического АГП.
+Лемма доказана.
+
+%Установим условия сходимости для АГП-Н и покажем, что расширение алгоритма вычисления характеристики не рождает предельных точек другого типа по сравнению с предельными точками классического АГП.
 %, кроме предельных точек областей невычислимости.
@@
 \]
-то любая точка глобального минимума $y^* \in Q$ является предельной точкой последовательности $\{y^k\}$ и любая точка $\overline{y}$ является либо точкой глобального минимума функции $\phi(y), y \in Q$, либо принадлежит области невычислимости $\overline{y} \in I$.
+то любая точка глобального минимума $y^* \in Q$ является предельной точкой последовательности $\{y^k\}$, а любая предельная точка $\overline{y}$ является либо точкой глобального минимума функции $\phi(y), y \in Q$, либо принадлежит области невычислимости, т.е.
+$\overline{y} \in I$.
 
-\textbf{Доказательство}.
-Во-первых, докажем, что для любой предельной точки $\overline{y}=y(\overline{x}), \overline{x} \in [0,1]$ существует бесконечная последовательность вложенных интервалов, соответсвующих упорядоченным по возрастанию номерам итераций $k_i$, $\{(x_{t-1}, x_t)\}, t = t(k_i), i = 1, 2, ...$, удовлетворяющая условиям:
+\textbf{Доказательство}. Проведем доказательство теоремы в три этапа.
+
+1.
+Докажем, что для любой предельной точки $\overline{y}=y(\overline{x}), \overline{x} \in [0,1]$ существует бесконечная последовательность вложенных интервалов, соответсвующих упорядоченным по возрастанию номерам итераций $k_i$, $\{(x_{t-1}, x_t)\}, t = t(k_i), i = 1, 2, ...$, удовлетворяющая условиям:
 \[
@@
 \[
-\overline{x}=\bigcap_{i = 1}^{\infty}{(x_{t-1}, x_t)},
+\overline{x}=\bigcap_{i = 1}^{\infty}{[x_{t-1}, x_t]},
 \]
@@
 \]
-где $\Delta_t = (x_t - x_{t-1})^{1/N}$, $R(t)$ и $t$ -- из Правила 5 раздела 3.1.
+где $\Delta_t = (x_t - x_{t-1})^{1/N}$, а $R(t)$ и $t$ -- из (\ref{Rt}).
 
@@
 в интервал, содержащий предельную точку $\overline{x}$, попадает точка очередного испытания $x^{k_{i}+1}$.
-Из правил вычисления новой точки (\ref{eq10})--(\ref{eq11}), а также оценки (\ref{eq7}) получаем
-оценку для максимальной длины очередного поискового интервала
+Из правил вычисления новой точки (\ref{eq11}), (\ref{eq10}), а также оценки (\ref{eq7}) получаем
+оценку для максимальной длины нового поискового интервала
 %\[
@@
 \]
-где коэффициент $\gamma = \frac{r+1}{2r} < 1$. Отсюда следует, что $\Delta_t \to 0$ при $i \to \infty$ и $\overline{x}=\bigcap_{i = 1}^{\infty}{(x_{t-1}, x_t)}$.
+где коэффициент $\gamma = \frac{r+1}{2r} < 1$. Отсюда следует, что $\Delta_t \to 0$ при $i \to \infty$ и $\overline{x}=\bigcap_{i = 1}^{\infty}{[x_{t-1}, x_t]}$.
 
 %В предположении, что на каждом шаге $k \geq 1$ существует некоторый интервал $(x_{j-1}, 
-Предположим, что на каждом шаге $k \geq h$ существует некоторый интервал $(x_{j-1}, x_{j})$, в одной из точек которого достигается текущее минимальное значение $z^*$, т.е.
+Рассмотрим теперь некоторый шаг поиска $k \geq 1$ и интервал $(x_{j-1}, x_{j})$, в одной из граничных точек которого достигается текущее минимальное значение $z^*$, т.е.
 $z^* = z_j$ или $z^* = z_{j-1}$.
-Тогда справедливо соотношение $z_j + z_{j-1}-2z^* = |z_j-z_{j-1}| = \mu_j \Delta_j \leq \mu \Delta_j$ и, если обе граничные точки вычислимы $v(x_{j-1}) = v(x_j) = 1$, то из (\ref{eq9}) следует
+Такой интервал будет существовать в силу доказанной ранее леммы.
+В случае, если $v(x_{j-1}) = v(x_j) = 1$, т.е.
+обе точки $y(x_{j-1})$ и $y(x_{j})$ принадлежат области вычислимости $Q$, то справедливо соотношение
 \[
-R(j) = \Delta_j + \frac{(\mu_j \Delta_j)^2}{(r\mu)^2 \Delta_j}-2\frac{\mu_j \Delta_j}{r\mu} = (\Delta_j-\frac{\mu_j\Delta_j}{r\mu})^2 \geq (1-1/r)^2 \Delta_j > 0,
+z_j + z_{j-1}-2z^* = |z_j-z_{j-1}| = \mu_j \Delta_j \leq \mu \Delta_j, 
 \]
-а если $v(x_{j-1}) \neq v(x_j)$, то $R(j)=2\Delta_j>0$.
-Таким образом, на каждом шаге поиска $k \geq h$ существует интервал со сторого положительной характеристикой.
+а из формулы (\ref{eq9}) следует неравенство
+\begin{equation}\label{Rlower} 
+R(j) = \Delta_j + \frac{(\mu_j \Delta_j)^2}{(r\mu)^2 \Delta_j}-2\frac{\mu_j \Delta_j}{r\mu} = (\Delta_j-\frac{\mu_j\Delta_j}{r\mu})^2 \geq (1-1/r)^2 \Delta_j > 0.
+\end{equation}
+В случае, если $v(x_{j-1}) \neq v(x_j)$, то $R(j)=2\Delta_j>0$. 
+%Если же попаданий точек испытаний в область вычислимости не было, т.е. $v(x_{j-1}) = v(x_j) = 0$, а значение  $z^*$ является неопределенным, то $R(j)=(1-\frac{1}{r})^2 \Delta_j > 0$. 
+Таким образом, на каждом шаге поиска при достаточно большом $k$ существует интервал со сторого положительной характеристикой.
 Значит и характеристики лучших интервалов с номерами $t = t(k_i)$ будут положительными, т.е.
@@
 \]
-В случае если $v(x_{t-1}) = v(x_t) = 0$ или $v(x_{t-1}) = 0$ и $v(x_t) = -1$ или $v(x_{t-1}) = -1$ и $v(x_t) = 0$, имеет место строгое равенство $R(t) = (1-1/r)^2\Delta_t$, в остальных случаях $R(t) \leq 2\Delta_t$.
-Из полученных оценок и в силу уже доказанного ранее свойства $\lim_{i \to \infty}{\Delta_t} = 0$ следует, что $\lim_{i \to \infty}{R(t(k_i))}=0, i = 1, 2, ...$.
+Для случая $v(x_{t-1}) = v(x_t) = 0$ имеет место строгое равенство $R(t) = \alpha (1-1/r)^2\Delta_t$, в остальных случаях $R(t) \leq 2\Delta_t$.
+Из полученных оценок и в силу уже доказанного ранее свойства $\lim_{t \to \infty}{\Delta_t} = 0$ следует, что $\lim_{i \to \infty}{R(t(k_i))}=0, i = 1, 2, ...$.
 
-Во-вторых, покажем, что если $y^* = y(x^*) \in Q$ -- точка глобального минимума, то $y^*$ является предельной точкой последовательности $\{y^k\}$.
+2. Докажем, что если $y^* = y(x^*) \in Q$ -- точка глобального минимума, то $y^*$ является предельной точкой последовательности $\{y^k\}$.
 
@@
 \]
-а в силу условия Гелдера (\ref{eq4}) справедливы оценки
+а в силу условия Гельдера (\ref{eq4}) справедливы оценки
 \[
@@
 \[
-R(j) =(1-\frac{1}{r})^2\Delta_i>0, \text{ если } v(x_j) = v(x_{j-1}) = 0,
+R(j) =\alpha \left(1-\frac{1}{r}\right)^2\Delta_i>0, \text{ если } v(x_j) = v(x_{j-1}) = 0,
 \]
@@
 \[
-\leq \Delta_j(1-\frac{2^{3-1/N}L\sqrt{N+3}}{r\mu})>0, \text{ если } v(x_j) = v(x_{j-1}),
+\leq \Delta_j(1-\frac{2^{3-1/N}L\sqrt{N+3}}{r\mu})>0, \text{ если } v(x_j) = v(x_{j-1}) = 1,
 \]
@@
 Значит в интервал $(x_{j-1}, x_{j})$ попадёт точка одного из следующих испытаний.
-Повторяя данные рассуждения, приходим к выводу, что правила алгоритма порождают вложенную последовательность сжимающихся к $x^*$ интервалов, т.е.
+Повторяя данные рассуждения, приходим к выводу, что правила алгоритма порождают вложенную последовательность стягивающихся к $x^*$ интервалов, т.е.
 $x^*$ -- предельная точка и $\lim_{k \to \infty}{\{x^k\}} = x^*$.
 
-В-третьих, покажем, что если $\overline{y} = y(\overline{x})$ -- предельная точка последовательности $\{y^k\}$, то $\overline{y}$ является либо точкой глобального минимума функции $\phi(y), y \in Q$ , либо принадлежит области невычислимости $\overline{y} \in I$.
+3.
+Докажем, что если $\overline{y} = y(\overline{x})$ -- предельная точка последовательности $\{y^k\}$, то $\overline{y}$ является либо точкой глобального минимума функции $\phi(y), y \in Q$, либо принадлежит области невычислимости, т.е.
+$\overline{y} \in I$.
 
@@
 \]
-Тогда для очередного лучшего интервала при достаточно больших t справедливо $z_{t-1} \geq z^* + \delta$, если $v(x_{t-1}) = 1$, и $z_{t} \geq z^* + \delta$, если $v(x_{t}) = 1$. 
-Из правил вычисления характеристик следует
+Тогда для очередного лучшего интервала при достаточно больших $t$ справедливо $z_{t-1} \geq z^* + \delta$, если $v(x_{t-1}) = 1$, и $z_{t} \geq z^* + \delta$, если $v(x_{t}) = 1$. 
+Из правил вычисления характеристик следует, что 
 \[
-R(t) < \Delta_t-\frac{4\delta}{r\mu} < 0, \text{ если } v(x_{t}) = 1 \text{ и } v(x_{t-1}) = 1,
+R(t) < \Delta_t-\frac{4\delta}{r\mu} < 0, \text{ если } v(x_{t-1}) = v(x_{t}) = 1,
 \]
 \[
-R(t) < 2\Delta_t-\frac{4\delta}{r\mu} < 0, \text{ если } v(x_{t}) \neq v(x_{t-1}),
+R(t) < 2\Delta_t-\frac{4\delta}{r\mu} < 0, \text{ если } v(x_{t}) \neq v(x_{t-1}).
 \]
-так как, как было показано ранее $\lim_{t \to \infty}{\Delta_t} = 0$, что противоречит с доказанным ранее свойством положительности харакетристик стягивающихся интервалов $R(t) > 0$.
-Значит наше предположение неверно и либо предельная точка $\overline{y} = y(\overline{x})$ является точкой глобального минимума, либо функция в ней является невычислимой и алгоритмом порождается последовательность невычислимых стягивающихся интервалов с положительной характеристикой, вычисляемой по правилу $R(t) =(1-\frac{1}{r})^2\Delta_t$.
+Ранее было доказана справедливость $\lim_{t \to \infty}{\Delta_t} = 0$, что противоречит свойству положительности харакетристик стягивающихся интервалов $R(t) > 0$.
+Значит, наше предположение неверно и либо предельная точка $\overline{y} = y(\overline{x})$ является точкой глобального минимума, либо функция в ней является невычислимой.
+%и алгоритмом порождается последовательность невычислимых стягивающихся интервалов с положительной характеристикой, вычисляемой по правилу $R(t) =(1-\frac{1}{r})^2\Delta_t$.
+
+Теорема доказана.
+
+\textbf{Замечание 1}.
+В случае пустой области невычислимости целевой функции, т.е.
+$I = \emptyset$ и $Q = D$, условия теоремы будут соответствовать теореме о сходимости исходного алгоритма глобального поиска.
+А именно, последовательность испытаний не будет иметь иных предельных точек, кроме точек глобального минимума целевой функции, и все точки глобального минимума будут являться предельными.
+
+\textbf{Замечание 2}.
+Неравенство (\ref{Rlower}), выведенное в ходе доказательства теоремы, поясняет способ вычисления характеристик поисковых интервалов $(x_{j-1}, x_{j})$, в обоих граничных точках которых значение функции является неопределенным.
+А именно, неравенство (\ref{Rlower}) дает нижнюю оценку для характеристики интервала, в граничной точке которого достигается текущее лучшее значение $z^*$.
+Основываясь на этой оценке, характеристику интервала, для которого $v(x_{j-1}) = v(x_{j}) = 0$, можно вычислять как $R(j) = \alpha {(1-\frac{1}{r})}^2 \Delta_j $.
+При этом параметр $ 0 \leq \alpha \leq 1$ можно интерпретировать как априорную оценку соотношения объемов областей $I$ и $Q$.
+Значение $\alpha << 1$ будет соответстовать предположению о малой области невычислимости целевой функции, тогда как значение $\alpha \approx 1$ соответствует предположению об области $I$, сравнимой по объему с областью $Q$.
 

### 2025-03-24 2025/RuSCDays/RSD-2025.tex
RuSCDays initial version

--- 2018-12-28 2019/MOTOR/motor.tex
@@
 %
-% If you use the hyperref package, please uncomment the following line
+% If you use the hyperref package, please uncomment the following two lines
 % to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
+%\usepackage{color}
+%\renewcommand\UrlFont{\color{blue}\rmfamily}
+%\urlstyle{rm}
+
+\usepackage{amsmath}
+\usepackage{marvosym}
+\usepackage{hyperref}
+
+\def\letter{$^{\textrm{(\Letter)}}$}
 
@@
 %
-\title{ Generalized nested optimization scheme for global optimization problems
-\thanks{This study was supported by the Russian Science Foundation, project 
-No.\,16-11-10150.}}
+\title{A Parallel Algorithm for Solving Global Optimization Problems and Its Application for Tuning Hyperparameters of AI Methods}
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{A Parallel Algorithm for Tuning Hyperparameters}
 % If the paper title is too long for the running head, you can set
@@
 %
-\author{Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
-Ilia Lebedev\orcidID{0000-0002-8736-0652}}
+\author{Marina Usova \letter \orcidID{0000-0002-0722-6884} \and
+Ilya Lebedev\orcidID{0000-0002-8736-0652}  \and
+Anton Shtanyuk\orcidID{0000-0003-1809-7173} \and
+Evgeny Kozinov \orcidID{0000-0001-6776-0096} \and
+Konstantin Barkalov \orcidID{0000-0001-5273-2471} 
+}
 %
-\authorrunning{K. Barkalov, I. Lebedev}
+\authorrunning{M. Usova et al.}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
-\email{konstantin.barkalov@itmm.unn.ru}}
+\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
+\email{usova@itmm.unn.ru}}
 %
@@
 %
-\section{First Section}
-\subsection{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is not indented. The first paragraph that follows a table, figure,
-equation etc. does not need an indent, either.
+
+\section{Introduction}
+Test links \cite{Seleznev2019,Blechschmidt2021,Xu2020,Strongin2000,Evtushenko2013,Evtushenko2009,Paulavicius2011,Gegrel2021,PaulaviciusZilinskas2014,Sergeyev2017,Usova2024,NIPS2015,Voevodin2015}
+
+\section{Problem Statement}
+Please note that the first paragraph of a section or subsection is not indented. The first paragraph that follows a table, figure, equation etc. does not need an indent, either.
 
@@
 
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
 \begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
+\caption{Table captions should be placed above the tables.}\label{tab1}
 \begin{tabular}{|l|l|l|}
@@
 
-\noindent Displayed equations are centered and set on a separate
-line.
+\noindent Displayed equations are centered and set on a separate line.
 \begin{equation}
@@
 \end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
+Please try to avoid rasterized images for line-art diagrams and schemas. Whenever possible, use vector graphics instead (see
 Fig.~\ref{fig1}).
 
-%\begin{figure}
-%\includegraphics[width=\textwidth]{fig1.eps}
-%\caption{A figure caption is always placed below the illustration.
-%Please note that short captions are centered, while long ones are
-%justified by the macro package automatically.} \label{fig1}
-%\end{figure}
+\begin{figure}
+\center
+\includegraphics[width=0.8\textwidth]{fig1.png}
+\caption{A figure caption.} \label{fig1}
+\end{figure}
 
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
+%\begin{theorem}
+%This is a sample theorem. The run-in heading is set in bold, while
+%the following text appears in italics. Definitions, lemmas,
+%propositions, and corollaries are styled the same way.
+%\end{theorem}
 %
@@
 \end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{globalizerSystem}, a
-book~\cite{strSergGO}, proceedings without editors~\cite{Jones2009}.
+
+\begin{credits}
+\subsubsection{\ackname} This study was funded by the ...
+
+\subsubsection{\discintname}
+The authors have no competing interests to declare that are relevant to the content of this article.
+\end{credits}
 %
@@
 %
-
 %\begin{thebibliography}{8}
@@
 %\bibitem{ref_url1}
-%LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-%Oct 2017
+%LNCS Homepage, \url{http://www.springer.com/lncs}, last accessed 2023/10/25
 %\end{thebibliography}

### 2025-04-04 2025/RuSCDays/RSD-2025.tex
add teoretical part of text in english

@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-150--250 words.
+%(рус.) В статье рассматривается параллельный алгоритм для решения задач глобальной оптимизации с частично определенной целевой функцией вида «черный ящик».
+Задачи такого вида возникают в процессе выбора оптимальных значений гиперпараметров методов машинного обучения и искусственного интеллекта.
+Алгоритм использует схему редукции размерности на основе кривых Пеано в сочетании с асинхронной схемой распараллеливания вычислений.
+Для демонстрации эффективности распараллеливания проведены вычислительные эксперименты в задаче настройки гиперпараметров методов искусственного интеллекта, используемых для предсказания значений временного ряда.
+The paper considers a parallel algorithm for solving global optimization problems with a partially defined objective function of the “black box” type.
+Such problems arise in the process of selecting optimal values of hyperpa-rameters used in machine learning and artificial intelligence methods.
+The al-gorithm employs a dimensionality reduction scheme based on Peano curves in combination with an asynchronous scheme of computation parallelization.
+To demonstrate the effectiveness of parallelization, computational experi-ments were conducted for the task of tuning the hyperparameters of artificial intelligence methods used to predict the values of a time series.
 
-\keywords{First keyword  \and Second keyword \and Another keyword.}
+\keywords{Global optimization  \and Multiextremal Functions \and Partially Defined Functions \and Hyperparameter Tuning \and  Time Series \and Artificial Intelligence.}
 \end{abstract}
@@
 \section{Introduction}
-Test links \cite{Seleznev2019,Blechschmidt2021,Xu2020,Strongin2000,Evtushenko2013,Evtushenko2009,Paulavicius2011,Gegrel2021,PaulaviciusZilinskas2014,Sergeyev2017,Usova2024,NIPS2015,Voevodin2015}
+
+%(рус.) В настоящее время методы машинного обучения и искусственного интеллекта используются во многих научных и прикладных областях.
+С их помощью успешно решаются задачи, которые еще недавно решались только классическими численными методами [ссылки].
+Currently, machine learning and artificial intelligence methods are used in many academic and applied fields.
+They are successfully applied to address problems that until recently were solved only by classical numerical methods [REFS].
+
+%(рус.) В большинстве методов искусственного интеллекта имеются гиперпараметры, от выбора конкретных значений которых может зависеть эффективность их работы.
+Например, на качество классификации объектов с помощью метода опорных векторов, которое традиционно измеряется метрикой $F_1$ [ссылка], существенным образом влияют два параметра метода: коэффициент регуляризации и коэффициент ядра.
+При этом значения по умолчанию, ориентированные на средний случай, не будут являться наилучшими для конкретного класса решаемых задач.
+Подбор указанных параметров может значительно повысить значение целевой метрики качества.
+Most artificial intelligence methods involve hyperparameters, and the choice of their specific values may determine their performance.
+For example, the quality of object classification using the support vector method, which is traditionally measured by the metric $F_1$ [REF], is significantly affected by two parameters of the method: the regularization coefficient and the kernel coefficient.
+However, the default values geared to the average case will not be the best for a particular class of problems to be solved.
+Appropriate selection of these parameters can significantly increase the value of the objective quality metric.
+
+%(рус.) Таким образом, возникает the hyperparameter optimization (HPO) problem, которая с математической точки зрения является задачей глобальной оптимизации.
+Типичным является случай наличия не только непрерывных, но и категориальных параметров, т.е.
+возникает необходимость использования методы глобальной оптимизации, способные решать задачи со смешанными параметрами.
+Thus, there arises the hyperparameter optimization (HPO) problem, which from the mathematical point of view is a global optimization problem.
+In this case, it is typical to have not only continuous but also categorical parameters (parameters having a pre-fixed set of values), i.e.
+there is a need to use global optimization methods capable of solving problems with mixed parameters.
+
+%(рус.) Традиционную сложность HPO problem составляет значительное время, требуемое для оценки критерия при конкретном сочетании параметров.
+Поэтому применение для данных задач метаэвристических алгоритмов, требующих большого числа поисковых испытаний, является ограниченным.
+Более перспективным здесь является использование детерминированных алгорит-мов липшицевой оптимизации [Евтушенко, Сергеев, Стронгин], которые превосходят эвристические методы по качеству работы [Сергеев].
+Для методов липшицевой глобальной оптимизации предложен ряд подходов к распараллеливанию, обеспечивающих эффективную работу на суперкомпьютерных системах [Евтушенко, Жилинскас, Гергель].
+The traditional complexity of HPO problems is due to the considerable time required to evaluate a criterion for a particular combination of parameters.
+Therefore, metaheuristic algorithms, which require a large number of search tri-als, can find only a limited use for such problems.
+A more promising approach is to use deterministic Lipschitz optimization algorithms [Evtushenko, Sergeev, Strongin], which are superior to heuristic methods in terms of performance quality [Sergeev].
+For Lipschitz global optimization methods, a number of paral-lelization approaches have been proposed to ensure efficient operation in super-computer systems [Evtushenko, Ziliskas, Gergel].
+
+%(рус.) Еще одной сложностью рассматриваемого класса задач HPO является наличие таких сочетаний значений гиперпараметров, при которых в некоторых (заранее не известных) подобластях области поиска исследуемый метод в принципе не работает.
+С такими значениями параметров невозможно корректно провести расчет и вычислить значение критерия эффективности.
+Данное явление можно интерпретировать как частичную определенность критерия в области поиска.
+В такой постановке HPO problem существенно усложняется, т.к.
+область допустимых сочетаний параметров является заранее неопределенной.
+One more difficulty of the class of HPO problems being considered is the pres-ence of such combinations of hyperparameter values for which the method under study does not work in principle in some (not known in advance) subdomains of the search domain.
+With such values of parameters it is impossible to perform the calculation correctly and calculate the value of the efficiency criterion.
+This phenomenon can be interpreted as partial definiteness of the criterion in the search domain.
+In such a formulation, the HPO problem becomes significantly more complicated, since the region of permissible combinations of parameters is undefined a priori.
+
+%(рус.) Данная статья продолжает развитие подхода к параллельному решению задач липшицевой глобальной оптимизации, предложенного в работах [наши работы].
+Новым элементом является сочетание способов распараллеливания и предложенной в [Усова] схемы работы с частично вычислимыми функция-ми.
+Основная часть публикации имеет следующую структуру.
+В Section 2 приведена математическая постановка решаемой задачи.
+Section 3 содержит описание параллельного алгоритма глобального поиска.
+В Section 4 описан способ решения задач со смешанными (дискретными и непрерывными) па-раметрами, а также обсуждается способ работы с функциями, определенны-ми не во всей области поиска.
+Section 5 содержит результаты вычислитель-ных экспериментов, демонстрирующих эффективность предлагаемого под-хода при решении задачи оптимизации гиперпараметров методов прогнози-рования значений временного ряда.
+This paper continues to develop the approach to solving Lipschitz global optimization problems in parallel as proposed in [REFS ON OUR WORKS].
+The new element is the combination of parallelization methods and the scheme for handling par-tially computable functions proposed in [Usova].
+The main part of the publica-tion has the following structure.
+Section 2 presents the mathematical formulation of the problem to be solved.
+Section 3 contains a description of the parallel global search algorithm.
+Section 4 describes the method for solving problems with mixed (discrete and continuous) parameters, and also discusses the method for handling functions that are not defined in the entire search domain.
+Section 5 gives the results of computational experiments demonstrating the effectiveness of the proposed approach in solving the problem of optimizing the hyperparameters of time series forecasting methods.
 
 \section{Problem Statement}
-Please note that the first paragraph of a section or subsection is not indented. The first paragraph that follows a table, figure, equation etc. does not need an indent, either.
+%(рус.) В начале рассмотрим задачу оптимизации в простой постановке, только с непрерывными параметрами. Пусть требуется найти минимум функции $\phi(y)$, определенной в области $D$
+First, consider the optimization problem in a simpler formulation, with only continuous parameters. Let it be required to find the minimum of the function $\phi(y)$ defined in the domain $D$
 
-Subsequent paragraphs, however, are indented.
+\begin{equation} \label{problem}
+\phi(y^*) = \min \left\{\phi(y): y \in D, g_i(y) \leq 0, 1 \leq i \leq m\right\}, 
+\end{equation}
+\[ D=\left\{ y \in R^N: a_i \leq y_i \leq b_i, \; 1 \leq i \leq N\right\}. \]
+
+%(рус.) Будем предполагать, что целевая функция $\phi(y)$ является многоэкстремальной удовлетворяет условию Липшица
+We will assume that the objective function $\phi(y)$ is multiextremal and satisfies the Lipschitz condition
+\begin{equation}\label{lipschitz} 
+| \phi (y_1)-\phi (y_2) | \leq L \| y_1-y_2 \|, \; y_1,y_2 \in D.
+\end{equation}
+
+%(рус.) Известным способом решения задач липшицевой глобальной оптимизации является их сведение (явное или неявное) к решению одномерных задач.
+Такое сведение может быть выполнено, например, на основе использования диагональных или симплексных разбиений области поиска [ссылки].
+A well-known way to solve Lipschitz global optimization problems is to reduce them (explicitly or implicitly) to the solution of one-dimensional problems.
+Such a reduction can be performed, for example, by using diagonal or simplex partitions of the search domain [REFS].
+
+%(рус.) В нашем исследовании мы используем схему редукции размерности на основе кривых Пеано y(x), однозначно и непрерывно отображающих отрезок [0,1] на N-мерный гиперкуб D [ссылки].
+В результате такой редукции многомерная задача глобальной оптимизации (1) сводится к одномерной задаче
+In our research, we use a dimensionality reduction scheme based on Peano curves $y(x)$ that uniquely and continuously map the segment $[0,1]$ to an $N$-dimensional hypercube $D$ [REFS].
+As a result of this reduction, the multidimensional global optimization problem (\ref{problem}) reduces to a one-dimensional problem
+
+\begin{equation} \label{reduced_problem}
+\min \left\{\phi(y(x))\right\} = \min \left\{\phi(y)\right\}, x \in [0,1],  y \in D.
+\end{equation}
+
+%(рус.) Использование кривых Пеано сохраняет ограниченность разностей функции.
+Если исходная функция φ(x) удовлетворяла условию Липшица, то редуцированная одномерная функция φ(y(x)) из (3) будет удовлетворять условию Гёльдера
+By using Peano curves, the boundedness of the function differences is preserved.
+If the original function $\phi(x)$ satisfied the Lipschitz condition, then the reduced one-dimensional function $\phi(y(x))$ from (\ref{reduced_problem}) will satisfy the Holder condition
+\begin{equation}\label{holder} 
+| \phi (y(x'))-\phi (y(x'')) | \leq H | x' - x''|^\frac{1}{N},
+\end{equation}
+%(рус.) где константа H определяется соотношением H=2L√(N+3), а L есть константа Липшица из (2). 
+where the constant $H$ is defined by the relation $H=2L\sqrt{N+3}$, and $L$ is the Lipschitz constant from (\ref{lipschitz}).
+
+%(рус.) Таким образом, проведение испытания в точке x^k  ∈ [0,1] подразумевает:
+	%(рус.) вычисление image y^k   = y(x^k ) in accordance with the mapping y(x);
+	%(рус.) вычисление значения функции z^k  =φ(y^k ).
+Thus, performing a trial at a point $x^k \in [0,1]$ involves:
+\begin{itemize}
+    \item calculating the image $y^k=y(x^k)$ in accordance with the mapping $y(x)$;
+    \item calculating the value of the function $z^k=\phi(y^k)$.
+\end{itemize}
+
+\begin{figure}
+\center
+\includegraphics[width=0.8\textwidth]{fig1.png}
+\caption{Graph of the one-dimensional function after dimensionality reduction.} \label{fig1}
+\end{figure}
+
+%(рус.) При численной реализации оптимизационных алгоритмов используется развертка, приближающая истинную кривую Пеано с заданной точностью, которая определяется желаемой точностью поиска глобального экстремума.
+В качестве примера на Fig.
+1 приведена одномерная функция, получающаяся в результате редукции двумерной задачи.
+График демонстрирует сложную структуру получающейся одномерной функции.
+In the numerical implementation of optimization algorithms, an evolvent is used to approximate the true Peano curve with a given accuracy, which is deter-mined by the desired accuracy of the search for the global extremum.
+As an example, Fig.~\ref{fig1} shows a one-dimensional function resulting from the reduction of a two-dimensional problem.
+The graph demonstrates the complex structure of the resulting one-dimensional function.
+
+\section{Parallel asynchronous global search algorithm}
+
+%(рус.) К настоящему времени разработаны различные подходы к организации параллельных вычислений для различных численных алгоритмов.
+Однако при решении задач многомерной многоэкстремальной оптимизации известные способы или не обеспечивают должного эффекта, или вообще не применимы.
+Например, один из основных способов организации параллельных вычислений состоит в разделении области решения задачи между процессами и параллельного решения задачи в этих подобластях.
+Но в случае решения задач глобальной оптимизации такой подход будет неэффективным, поскольку при разделении области поиска только небольшая часть процессов будет работать в подобласти с искомым глобальным минимумом, все же остальные будут работать в подобластях, в которых отсутствует глобальный минимум исходной задачи.
+Various approaches to organizing parallel computations for various numerical algorithms have been developed to date [REFS].
+However, when solving multidimensional multiextremal optimization problems the known methods either do not provide proper effect or are not applicable altogether.
+For example, one of the main ways of organizing parallel computations consists in dividing the prob-lem solving domain between processes and solving the problem in parallel in these subdomains.
+However, this approach will be inefficient for solving global optimization problems, because when dividing the search domain, only a small part of the processes will run in the subdomain with the desired global minimum, while the rest will run in the subdomains in which there is no global minimum of the original problem.
+
+%(рус.) Более продуктивный подход к распараллеливанию основан на изменении схемы последовательного алгоритма для параллельного выполнения нескольких испытаний.
+Отметим, что в задачах рассматриваемого класса (задачах настройки гиперпараметров методов машинного обучения) типичной является ситуация, когда целевая функция имеет разную трудоемкость вычисления в разных точках области поиска.
+В таких задачах распределение точек испытаний между параллельно работающими процессами должно происходить асинхронно, по запросу.
+В самом деле, в случае синхронного проведения испытаний часть процессов (которые проводят испытаний в «легкой» подобласти) закончит вычисления и будет простаивать, дожидаясь окончания расчетов остальных процессов (которым достались точки из «сложной» подобласти).
+A more productive approach to parallelization is based on changing the sequential algorithm scheme for performing several trials in parallel.
+Note that in the problems of the class under consideration (problems of hyperparameter tun-ing for machine learning methods) a typical situation is when the objective function has different computational complexity at different points of the search domain.
+In such problems, the distribution of trial points between parallel processes should be performed asynchronously, on demand.
+Actually, in case of synchronous trials, some processes (which perform trials in the ``easy'' subdomain) will finish calculations and will idle, waiting for the rest of the processes (which were assigned the points from the ``difficult'' subdomain) to finish calculations.
+
+%(рус.) В асинхронном алгоритме глобального поиска реализована схема распараллеливания вида ''master/worker''.
+При описании алгоритма будем предполагать, что в нашем распоряжении имеется p+1 вычислительный процесс: один процесс-мастер и p рабочих процессов.
+Процесс-мастер организует сбор, накопление и обработку поисковой информации (результатов выполненных испытаний).
+На ее основе мастер вычисляет точки новых испытаний и распределяет их по процессам-рабочим.
+Процессы-рабочие получают от мастера точки, проводят в них новые испытания и отсылают мастеру их результаты.
+We used in our study an asynchronous global search algorithm, which implements a parallelizing scheme of the ``master/worker'' type.
+When describing the algorithm we will assume that we have at our disposal the following computational processp+1 : one master process andp worker processes.
+The master process organizes the collection, accumulation and processing of search information (results of performed trials).
+Based on this information, the master calculates the points of new trials and distributes them to the worker processes.
+The worker processes receive the points from the master, perform new trials in these points and send their results to the master.
+
+%(рус.) На первой итерации алгоритма процесс-мастер (равномерно или случайно) выбирает p различных внутренних точек области поиска {x^1,x^2,…,x^p }, x^i∈(0,1),i=1,…,p, и рассылает точки испытаний по процессам-рабочим для выполнения испытаний в них.
+Правила работы процесса-мастера на любой из следующих итераций состоят в следующем.
+At the first iteration of the algorithm, the master process selects (uniformly or randomly) p different internal points of the search domain ${x^1, x^2, ..., x^p}, x^i \in (0,1), i=1, ..., p$, and distributes the trial points to the worker processes to perform trials in those points.
+The rules for the master process at any of the subsequent iterations are as follows.
+
+%(рус.) Пусть проведено k испытаний, их результаты получены и обработаны мастером, а на текущей итерации процессы-рабочие проводят испытания в точках множества S_k  = {x^(k+1),x^(k+2),…,x^(k+p) }, т.е.
+в этих точках испытания уже начались, но еще не завершены.
+Как только один из процессов-рабочих завершает проведение испытания в некоторой точке (например, x^(k+1)), он пересылает процессу-мастеру результаты испытания.
+Получив от рабочего процесса результаты испытания и обработав их мастер, в свою очередь, выбирает для него точку нового испытания x^(k+p+1) в соответствии со следующими правилами.
+Suppose $k$ tests have been performed, the master has received and processed their results, and during the current iteration the worker processes perform trials at the points of the set $S_k={x^(k+1), x^(k+2), …, x^(k+p)}$, i.e.
+the trials have already started at these points but have not yet been completed.
+As soon as one of the worker processes completes a trial at some point (e.g.
+$x^(k+1)$), it forwards the results of the trial to the master process.
+After receiving the trial results from the worker process and processing them, the master process in turn selects a new trial point $x^(k+p+1)$ for the worker process according to the following rules.
+
+\begin{enumerate}
+%(рус.) 1. Перенумеровать в порядке возрастания все точки X_k  = {x^1,x^2,…,x^(k+p)  }, в которых либо проведены, либо проводятся испытания, т.е.
+\item Renumber, in ascending order, all the points $X_k={x^1,x^2,…,x^(k+p) }$ at which trials have either been carried out or are being carried out, i.e.
+\[
+0 = x_1 \leq x_2 \leq ... \leq x_{x+p} = 1,
+\]
+%(рус.) и сопоставить им значения z_i=φ(y(x_i )) , 1≤i≤k, вычисленные в этих точках, а также индекс v_i=v(x_i ), определяемый по правилу
+and assign to them the values $z_i = \phi(y(x_i)), 1 \leq i \leq k$ calculated at these points, as well as the index $v_i = v(x_i)$ determined by the rule
+
+\begin{equation}\label{rule_points_index} 
+v_i=v(x_i)=
+\begin{cases}
+    0     & \quad \text{if } x_i \text{ is boundary point},\\
+    1     & \quad \text{if } x_i \text{ is internal point}.
+\end{cases}
+\end{equation}
+%(рус.) Точки x_0=0 и x_(k+p)=1 введены дополнительно (значения z_0 и z_(k+p) не определены) для удобства последующих обозначений.
+The points $x_0=0$ and $x_{k+p}=1$ are additionally introduced (the values $z_0$ and $z_{k+p}$ are not defined) for the convenience of subsequent designations.
+
+%(рус.) 2. Вычислить значения
+\item Calculate the values
+\[
+M_1= \max \left\{ \frac{|z_i - z_{i - 1}|}{(x_i - x_{i - 1})^\frac{1}{N}}: x_{i-1} \notin S_k, x_i \notin S_k, 2 \leq i \leq k+p \right\},
+\]
+\[
+M_2= \max \left\{ \frac{|z_{i+1} - z_{i - 1}|}{(x_{i+1} - x_{i - 1})^\frac{1}{N}}: x_{i} \in S_k, 2 \leq i < k+p \right\},
+\]
+\[
+M= \max \left\{ M_1, M_2 \right\},
+\]
+%(рус.) где z_i=f(x_i ), if x_i  ∉  S_k,1≤ i≤ k+p.
+Значения z_i в точках x_i∈S_k являются неопределенными, т.к.
+испытания в этих точках еще не завершены.
+Если значение M получилось равным 0, то положить M=1.
+where $z_i=f(x_i)$, if $x_i \notin S_k, 1 \leq i \leq k+p$.
+The values $z_i$ at the points $x_i \in S_k$ are undefined because the trials at these points have not yet been completed.
+If the value $M$ turns out to be equal to $0$, then assume $M=1$.
+
+%(рус.) 3. Определить текущее лучшее значение целевой функции
+\item Determine the current best value of the objective function
+\begin{equation}\label{cur_best_val} 
+z^* = \min \left\{ \phi(x_i): 1 \leq i \leq k \right\}.
+\end{equation}
+
+%(рус.) 4. Каждому интервалу (x_(i-1),x_i ),x_(i-1)  ∉ S_k,x_i  ∉ S_k,2≤ i≤ k+p, поставить в соответствие число R(i), которое называется характеристикой интервала и вычисляется по формуле
+\item Assign a number $R(i)$ to each interval $(x_{i-1}, x_i), x_(i-1) \notin S_k, x_i \notin S_k, 2 \leq i \leq k+p$, which is called the interval characteristic and is calculated using the formula
+\begin{equation}\label{characteristic}
+R(i)=
+\begin{cases}
+    \Delta_i+\frac{(z_i-z_{i-1})^2}{(r\mu)^2\Delta_i} - 2 \frac{(z_i+z_{i-1}-2z^*)}{r\mu},   & \quad v(x_i)=v(x_{i-1})=1,\\
+    2\Delta_i- 4 \frac{(z_i-z^*)}{r\mu},   & \quad v(x_i)=1, v(x_{i-1})=0,\\
+    2\Delta_i- 4 \frac{(z_{i-1}-z^*)}{r\mu},       & \quad v(x_i)=0, v(x_{i-1})=1,\\
+\end{cases}
+\end{equation}
+%(рус.) где Δ_i  =(x_i-x_(i-1) )^(1/N) , and r>1 – параметр алгоритма.
+where $\Delta_i=(x_i-x_{i-1})^\frac{1}{N}$, and $r>1$ is the parameter of the algorithm.
+
+%(рус.) 5. Выбрать интервал [x_(t-1),x_t ], которому соответствует максимальная характеристика, т.е.
+\item Choose the interval $[x_{t-1}, x_t]$ to which the maximum characteristic corresponds, i.e.
+\[
+R(t)= \max \left\{ R(i): x_{i-1} \notin S_k, x_i \notin S_k, 2 \leq i \leq k+p\right\}.
+\]
+
+%(рус.) 6. Вычислить точку проведения нового испытаний x^{k+p+1} ∈ (x_(t-1),x_t ) в соответствии с формулами
+\item Calculate the point of the new trial $x^{k+p+1} \in (x_{t-1}, x_t)$ according to the formulas
+\begin{equation}\label{new_point_1} 
+x^{k+p+1} = \frac{x_t+x_{t-1}}{2}, \text{ if } v(x_{t-1}) \neq v(x_t),
+\end{equation}
+\begin{equation}\label{new_point_2} 
+x^{k+p+1} = \frac{x_t+x_{t-1}}{2}-\textit{sign}{(z_t-z_{t-1})}\frac{|z_t-z_{t-1}|^N}{2r\mu^N}, \text{ if } v(x_{t-1})=v(x_t).
+\end{equation}
+%(рус.) Сразу после вычисления точки очередного испытания процесс-мастер добавляет ее в множество S и пересылает процессу-рабочему, который начинает проведение испытания в ней.
+Immediately after calculating the point of the next trial, the master process adds it to the set $S$ and forwards it to the worker process, which starts a trial there.
+\end{enumerate}
+
+%(рус.) Работа алгоритма завершается, если в мастер-процессе выполняется одно из двух условий: Δ_t<ϵ or k+p>K_max.
+Выполнение первого неравенства соответствует остановке алгоритма по заданной точности поиска 0<ϵ<1, выполнение второго – по заданному числу испытаний K_max>0.
+The algorithm is terminated if one of two conditions is met in the master process: $\Delta_t< \varepsilon$ or $k+p>K_{max}$.
+The fulfillment of the first inequality corresponds to stopping the algorithm for a given search accuracy $0< \varepsilon <1$, and the fulfillment of the second one corresponds to a given number of trials $K_{max}>0$.
+
+%(рус.) Теоретические основы алгоритма, а также примеры его применения приведены в работах [ссылки].
+The theoretical background of the algorithm as well as examples of its application are given in  [REFS].
+
+\section{Handling mixed parameters and functions that are not universally defined}
+
+%(рус.) Теперь рассмотрим задачу оптимизации в более сложной постановке. Пусть целевая функция задачи зависит от двух векторов параметров: вектора y∈D,  и вектора u∈U, т.е.
+Now consider the optimization problem in a more complex setting. Let the objective function of the problem depend on two vectors of parameters: vector $y \in D$, and vector $u \in U$, i.e.
+
+\begin{equation}\label{mixed_problem} 
+\phi(y^*, u^*) = \min \left\{\phi(y, u): y \in D, u \in U\right\}.
+\end{equation}
+
+%(рус.) Задача оптимизации в такой постановке является математической моделью проблемы поиска наилучшей комбинации гиперпараметров методов искусственного интеллекта.
+Здесь D соответствует области изменения непрерывных гиперпараметров, а множество U содержит возможные сочетания категориальных гиперпараметров.
+The optimization problem in this formulation is a mathematical model of the problem of finding the best combination of hyperparameters of artificial intelligence methods.
+Here, $D$ corresponds to the region of variation of continuous hyperparameters, and the set $U$ contains possible combinations of categorical hyperparameters.
+
+%(рус.) Задачу со смешанными параметрами можно свести к решению взаимосвязанного набора задач с непрерывными параметрами.
+Занумеруем целыми числами s все различные комбинации категориальных параметров, т.е.
+сопоставим каждому номеру s вектор u_s.
+Тогда (10) можно переписать в виде
+A problem with mixed parameters can be reduced to solving an interrelated set of problems with continuous parameters.
+Let us number all different combinations of categorical parameters with integers $s$, i.e., let us assign to each number $s$ a vector $u_s$.
+Then (\ref{mixed_problem} ) can be rewritten as
+
+\begin{equation}\label{mixed_problem_2} 
+ \min \left\{\phi(y, u)\right\} = \min_{s \in \{1, ..., s\}} \left\{\phi(y, u_s): y \in D\right\}.
+\end{equation}
+
+%(рус.) Используя схему редукции размерности с помощью кривых Пеано y(x), каждой задаче минимизации из (11) можно сопоставить одномерную задачу минимизации, т.е.
+сформировать множество одномерных задач
+Using the dimensionality reduction scheme with Peano curves $y(x)$, each minimization problem from (\ref{mixed_problem_2}) can be matched with a one-dimensional minimization problem, i.e.
+we can obtain a set of one-dimensional problems
+\begin{equation}\label{reduced_mixed_problem} 
+ \min \left\{\phi(y(x), u_s): x \in [0,1]\right\}, s \in \{1, ..., s\}.
+\end{equation}
+
+%(рус.) Если теперь ввести в рассмотрение отображение
+If we now introduce the mapping
+\[
+Y(x) = y(x-E(x)), x \in [0, S],
+\]
+%(рус.) переводящее любую точку интервала [0,S] на область D (обозначение E(x) соответствует целой части числа x) и определить функцию
+which translates any point of the interval $[0,S]$ to the domain $D$ (the notation $E(x)$ corresponds to the integer part of the number $x$) and define the function
+\[
+f(x)=\phi(Y(x), uE(x) + 1), x \in [0, S],
+\]
+%(рус.) можно переформулировать задачу (11) как
+we can reformulate problem (\ref{reduced_mixed_problem} ) as
+\begin{equation}\label{reduced_mixed_problem_2} 
+ \min \left\{f(x): x \in [0,S]\right\}.
+\end{equation}
+
+%(рус.) Применяя к решению задачи (12) алгоритм глобального поиска, мы найдем решение задачи (10). 
+Applying the global search algorithm to solve problem (\ref{reduced_mixed_problem_2}), we will find a so-lution to problem (\ref{mixed_problem}).
+
+%(рус.) Отметим, что функция f(x), вообще говоря, может иметь разрывы в целочисленных точках x_i  = i.
+Так как эти точки известны заранее, их можно на первой итерации добавить в информационную базу алгоритма со значением индекса равным 0.
+В соответствии с правилами алгоритма, в процессе его работы не будет возникать интервалов, обе граничные точки которых имеют нулевой индекс, и правила вычисления характеристик (7) и точки нового интервала (8)-(9) будут валидными.
+Note that the function $f(x)$, in general, can have discontinuities at integer points $x_i=i$.
+Since these points are known in advance, they can be added to the information base of the algorithm at the first iteration with an index value equal to $0$.
+According to the rules of the algorithm, during its operation there will be no intervals with both boundary points having zero index, and the rules for calculating the features (\ref{characteristic}) and the points of the new interval (\ref{new_point_1})-(\ref{new_point_2}) will be valid.
+
+%(рус.) Однако в рассматриваемых задачах настройки гиперпарметров целевая функция может быть не определена в некоторых (заранее неизвестных) точ-ках области поиска.
+Это может приводить возникновению интервалов, у ко-торых обе граничных точки имеют нулевой индекс, т.е.
+значение целевой функции в них не определено.
+Возможный подход к обработке таких интер-валов состоит в использовании на шаге 4 алгоритма для вычисления характе-ристики вместо формулы (7) следующей формулы
+However, in the hyperparameter tuning tasks under consideration, the objective function may be undefined in some (previously unknown) points in the search domain.
+This may give rise to intervals where both boundary points have a zero index, i.e.
+the value of the objective function is undefined in these points.
+A possible approach to processing such intervals is to use the following formula at step 4 of the algorithm to calculate the characteristic instead of formula (\ref{characteristic})
+
+\begin{equation}\label{characteristic_2}
+R(i)=
+\begin{cases}
+    \Delta_i+\frac{(z_i-z_{i-1})^2}{(r\mu)^2\Delta_i} - 2 \frac{(z_i+z_{i-1}-2z^*)}{r\mu},   & \quad v(x_i)=v(x_{i-1})=1,\\
+    2\Delta_i- 4 \frac{(z_i-z^*)}{r\mu},   & \quad v(x_i)=1, v(x_{i-1})=0,\\
+    2\Delta_i- 4 \frac{(z_{i-1}-z^*)}{r\mu},       & \quad v(x_i)=0, v(x_{i-1})=1,\\
+    \alpha(1-1/r)^2\Delta_i,       & \quad v(x_i)=0, v(x_{i-1})=0,\\
+\end{cases}
+\end{equation}
+
+%(рус.) В последнем выражении 〖 (1-1\/r)〗^2 ∆_i соответствует нижней оценке характеристики интервала, содержащего текущее найденное решение z^* из (6) (вывод данной оценки см.
+в [ссылка]).
+А параметр 0<α<1 соответствует априорный оценке доли объема области поиска, в которой значения функции являются неопределенными.
+In the latter expression, $\alpha(1-1/r)^2\Delta_i$ corresponds to the lower estimate of the characteristic of the interval containing the current solution $z^*$ found from (\ref{cur_best_val}) (see [REF]) for the derivation of this estimate).
+The parameter $0<\alpha<1$ corresponds to an a priori estimate of the portion of the search domain volume in which the function values are undefined.
+
+%(рус.) Также требуется дополнить правило выбора точки нового испытания для обработки ситуаций, когда обе граничные точки имеют нулевой индекс:
+It is also necessary to supplement the rule for selecting a new trial point to handle situations where both boundary points have a zero index:
+\begin{equation}\label{new_point_1_2} 
+x^{k+p+1} = \frac{x_t+x_{t-1}}{2}, \text{ if } v(x_{t-1}) \neq v(x_t) \text{ or } v(x_{t-1})=v(x_t)=0,
+\end{equation}
+\begin{equation}\label{new_point_2_2} 
+x^{k+p+1} = \frac{x_t+x_{t-1}}{2}-\textit{sign}{(z_t-z_{t-1})}\frac{|z_t-z_{t-1}|^N}{2r\mu^N}, \text{ if } v(x_{t-1})=v(x_t) = 1.
+\end{equation}
+
+%(рус.) Формула (14) будет применяться, если значения как минимум в одной из граничных точек интервала значения целевой функции не определены, тогда как формула (15) будет использоваться только в случае, когда значения це-левой функции вычислены в обоих точках.
+Formula (\ref{new_point_1_2}) will be used if the values in at least one of the boundary points of the objective function value interval are undefined, while formula (\ref{new_point_2_2}) will be used only when the values of the objective function are calculated at both points.
+
+\section{Results of computational experiments}
+
+Test links \cite{Seleznev2019,Blechschmidt2021,Xu2020,Strongin2000,Evtushenko2013,Evtushenko2009,Paulavicius2011,Gegrel2021,PaulaviciusZilinskas2014,Sergeyev2017,Usova2024,NIPS2015,Voevodin2015}
 
@@
 Fig.~\ref{fig1}).
-
-\begin{figure}
-\center
-\includegraphics[width=0.8\textwidth]{fig1.png}
-\caption{A figure caption.} \label{fig1}
-\end{figure}
 

### 2025-04-06 2025/RSCD_MCO/RSD-2025.tex
MCO список литературы + заголовок

--- 2025-03-24 2025/RuSCDays/RSD-2025.tex
@@
 %
-\title{A Parallel Algorithm for Solving Global Optimization Problems and Its Application for Tuning Hyperparameters of AI Methods}
+\title{Parallel algorithm for solving multicriterial optimization problems using elements of machine learning}
 %
-\titlerunning{A Parallel Algorithm for Tuning Hyperparameters}
+\titlerunning{Parallel algorithm for solving multicriterial optimization problems}
 % If the paper title is too long for the running head, you can set
@@
 %
-\author{Marina Usova \letter \orcidID{0000-0002-0722-6884} \and
-Ilya Lebedev\orcidID{0000-0002-8736-0652}  \and
-Anton Shtanyuk\orcidID{0000-0003-1809-7173} \and
-Evgeny Kozinov \orcidID{0000-0001-6776-0096} \and
-Konstantin Barkalov \orcidID{0000-0001-5273-2471} 
+\author{{Sergey Konnov\orcidID{0009-0003-4590-0870} \and
+Evgeny Kozinov \Letter\orcidID{0000-0001-6776-0096} \and
+Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
+Alexander Sysoyev \orcidID{0000-0003-1542-7624} \and
+Vladimir Grishagin\orcidID{0000-0002-2884-3670}}
 }
 %
-\authorrunning{M. Usova et al.}
+\authorrunning{S. Konnov, E. Kozinov, K. Barkalov, A. Sysoyev and V. Grishagin}
 % First names are abbreviated in the running head.
@@
 \institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
-\email{usova@itmm.unn.ru}}
+\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, \email{vagris@unn.ru}}
 %
@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-150--250 words.
+Solving problems of multi-criteria optimization involves finding many combinations of optimization parameters, in which the values of the criteria cannot be improved in several criteria at once without worsening the values   of the remaining ones (Pareto set).
+In the framework of the research, several assumptions were introduced: criteria are multiextremal and difficult to calculate, presented in the form of a ``black box'', and the number of optimized parameters is small.
+The paper presents a method for solving this class of problems based on the information-statistical approach and a machine learning procedure used to increase search efficiency.
+A parallel implementation of the method is described, which is supplemented with the ability to perform calculations in asynchronous parallel mode.
+The efficiency and scalability of the proposed algorithm is analyzed on the base of solving a test class of multiextremal problems.
 
-\keywords{First keyword  \and Second keyword \and Another keyword.}
+\keywords{Multicriterial Problems \and Global Optimization \and Machine Learning \and High Performance Computing.}
 \end{abstract}
@@
 \section{Introduction}
-Test links \cite{Seleznev2019,Blechschmidt2021,Xu2020,Strongin2000,Evtushenko2013,Evtushenko2009,Paulavicius2011,Gegrel2021,PaulaviciusZilinskas2014,Sergeyev2017,Usova2024,NIPS2015,Voevodin2015}
 
-\section{Problem Statement}
-Please note that the first paragraph of a section or subsection is not indented. The first paragraph that follows a table, figure, equation etc. does not need an indent, either.
-
-Subsequent paragraphs, however, are indented.
-
-\begin{table}
-\caption{Table captions should be placed above the tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
-
-\begin{figure}
-\center
-\includegraphics[width=0.8\textwidth]{fig1.png}
-\caption{A figure caption.} \label{fig1}
-\end{figure}
-
-%\begin{theorem}
-%This is a sample theorem. The run-in heading is set in bold, while
-%the following text appears in italics. Definitions, lemmas,
-%propositions, and corollaries are styled the same way.
-%\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
+\cite{Miettinen1999,Ehrgott2005,Pardalos2017,Konnov2025,Evtushenko2014,DPA02,Durillo2010,Mostaghim2007,Nebro2009,RC05,Zitzler2001,Gergel2019_2,Gergel2018,GergelKozinov2020,Marler2004,Strongin2000,Sergeyev2013,MCO_ML_2023,scikit-learn,PyTorch,ioptmco,HV,pymoo,AGP_ML}
 

### 2025-04-06 2025/RSCD_MCO/RSD-2025.tex
MCO 2025 рисунки

@@
 
+\begin{figure}
+\center
+\includegraphics[width=0.6\textwidth]{fig1.png}
+\caption{Computational scheme of the algorithm for solving multicriterial optimization problems using machine learning models.} \label{fig1}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width=0.6\textwidth]{fig2.png}
+\caption{Example of Pareto set evaluation in solving two scalar optimization problems.} \label{fig2}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width=0.8\textwidth]{fig3.png}
+\caption{Neural network models used to calculate the characteristic $R_{PS}(i)$.} \label{fig3}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width=0.6\textwidth]{fig4.png}
+\caption{Examples of Pareto set evaluation for the case of two scalar optimization problems.} \label{fig4}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width=0.8\textwidth]{fig5.png}
+\caption{Comparison of HW index values for sequential algorithm implementation.} \label{fig5}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width=0.8\textwidth]{fig6.png}
+\caption{Dependence of the mean value of the index HW from the trial number per process for the case of 20 processes.} \label{fig6}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width=0.8\textwidth]{fig7.png}
+\caption{Dependence of the HW index on the number of trials executed for different number of processes.} \label{fig7}
+\end{figure}
+
 \begin{credits}

### 2025-04-06 2025/RSCD_MCO/RSD-2025.tex
MCO 2025 Структура + таблицы

@@
 \section{Introduction}
+\label{sec1}
 
 \cite{Miettinen1999,Ehrgott2005,Pardalos2017,Konnov2025,Evtushenko2014,DPA02,Durillo2010,Mostaghim2007,Nebro2009,RC05,Zitzler2001,Gergel2019_2,Gergel2018,GergelKozinov2020,Marler2004,Strongin2000,Sergeyev2013,MCO_ML_2023,scikit-learn,PyTorch,ioptmco,HV,pymoo,AGP_ML}
+
+\section{Problem statement}
+\label{sec2}
+
+\section{Computational scheme of the developed algorithm}
+\label{sec3}
 
@@
 
+\subsection{Block of the global search algorithm}
+
+\subsection{Block for accumulating and reusing the information about current state of the search}
+
+\subsection{Block for building the Pareto set estimation}
+
 \begin{figure}
@@
 \end{figure}
+
+\subsection{Block of  building the machine learning for estimation of the Pareto set}
 
@@
 
+\subsection{Block for the choice of trial points and implementation of computational experiments}
+
+\section{Results of computational experiments}
+\label{sec4}
+
 \begin{figure}
@@
 \end{figure}
+
+% Please add the following required packages to your document preamble:
+% \usepackage{graphicx}
+\begin{table}[ht]
+\centering
+\caption{Average number of trials per process}
+\label{tab:1}
+\resizebox{0.8\textwidth}{!}{%
+\begin{tabular}{ccccccc}
+\hline
+\textbf{Number of processes}            & \textbf{1}      & \textbf{2}      & \textbf{4}      & \textbf{8}      & \textbf{16}    & \textbf{20}    \\ \hline
+MGSA                                    & 2269.82         & 1385.02         & 746.57          & 373.36          & 202            & 143.52         \\
+\textbf{MGSA   Linear SVC,  a = 0.05}   & \textbf{686.13} & \textbf{408.79} & \textbf{213.6}  & \textbf{111.85} & \textbf{70.95} & \textbf{62.16} \\
+MGSA   Poly SVC, a = 0.05               & 909.59          & 544.19          & 278.62          & 153.64          & 95.37          & 77.94          \\
+MGSA   RBF SVC, a = 0.04                & 652.74          & 446.84          & 233.81          & 119.84          & 72.36          & 62.37          \\
+\textbf{MGSA   Dist, a = 0.01}          & \textbf{659.98} & \textbf{402.9}  & \textbf{211.33} & \textbf{112.38} & \textbf{70.79} & \textbf{62.00} \\
+MGSA   NN 1 layer, a = 0.05             & 795.16          & 446.25          & 232.08          & 117.16          & 75.19          & 62.70          \\
+\textbf{MGSA   NN 2 layers, a =   0.03} & \textbf{682.12} & \textbf{401.74} & \textbf{219.62} & \textbf{113.32} & \textbf{73.51} & \textbf{61.32} \\ \hline
+\end{tabular}%
+}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Average HW Index in dependence on number of processes}
+\label{tab:2}
+\resizebox{0.8\textwidth}{!}{%
+\begin{tabular}{ccccccc}
+\hline
+Number of processes                   & \textbf{1}      & \textbf{2}      & \textbf{4}      & \textbf{8}      & \textbf{16}     & \textbf{20}     \\ \hline
+\textbf{MGSA}                         & \textbf{114.91} & \textbf{114.79} & \textbf{114.84} & \textbf{114.90} & \textbf{115.00} & \textbf{114.96} \\
+MGSA   Linear SVC, a = 0.05           & 114.44          & 114.65          & 114.72          & 114.74          & 114.91          & 114.96          \\
+\textbf{MGSA   Poly SVC,    a = 0.05} & \textbf{114.61} & \textbf{114.64} & \textbf{114.71} & \textbf{114.8}  & \textbf{114.95} & \textbf{115.01} \\
+MGSA   RBF SVC,    a = 0.04           & 114.54          & 114.54          & 114.73          & 114.89          & 114.92          & 114.96          \\
+MGSA   Dist, a = 0.01                 & 114.66          & 114.55          & 114.69          & 114.73          & 114.92          & 114.93          \\
+MGSA   NN 1 layer, a = 0.05           & 114.77          & 114.61          & 114.72          & 114.81          & 114.96          & 114.98          \\
+\textbf{MGSA   NN 2 layers, a = 0.03} & \textbf{114.74} & \textbf{114.54} & \textbf{114.71} & \textbf{114.76} & \textbf{114.98} & \textbf{115.01} \\ \hline
+\end{tabular}%
+}
+\end{table}
 
@@
 \end{figure}
+
+\section{Conclusion}
+\label{sec5}
+The research presents a parallel modification of the global search algorithm for solving multicriterial optimization problems.
+One of the features of the algo-rithm is the combination of the classic global search algorithm with machine learning methods.
+The combination of algorithms shows high efficiency in both sequential and parallel modifications.
+As the direction of further research it is planned to organize a number of experiments to solve applied problems and compare search efficiency with a number of third-party frameworks.
 

### 2025-04-06 2025/RSCD_MCO/RSD-2025.tex
MCO 2026 sec 1-3

@@
 
+The problems of choosing globally optimal solutions arise when conducting a wide class of researches.
+To find the best options, an efficiency criterion is determined, the values   of which depend on the parameters from the permissible domain.
+The criterion itself can often be presented not in an analytical form, but only as a procedure for calculating its values, i.e.
+in the form of a ``black box'' and additionally it can be multiextremal \cite{Miettinen1999,Ehrgott2005,Pardalos2017,Strongin2000,Sergeyev2013}.
+For a more accurate description of the problem being solved, several (generally speaking, contradictory) optimized criteria can be taken.
+
+Multiextremal problems of global optimization are computationally complicated because the globally optimal solution, because in the general case, to claim that the solution is globally optimal, you need to compare it with the values of the objective function at all other points in the search area.
+When solving such problems using the scanning methods based on random or uniform grids, the number of calculations increases exponentially with the growth of dimensionality.
+It is possible to solve the considered problems using more efficient algorithms that can be divided into two large classes: deterministic \cite{Miettinen1999,Ehrgott2005,Pardalos2017,Konnov2025,Evtushenko2014} and non-deterministic (meta-heuristic) methods \cite{DPA02,Durillo2010,Mostaghim2007,Nebro2009,RC05,Zitzler2001}.
+
+Non-deterministic algorithms often attempt to mimic the behavior of living organisms or processes occurring in nature \cite{DPA02,Durillo2010,Mostaghim2007,Nebro2009,RC05,Zitzler2001}.
+Algorithms from this class do not guarantee finding the best solution with a given accuracy, but can show good efficiency when solving problems with not too many number of local extrema.
+
+Deterministic algorithms guarantee an estimate of the globally optimal solution with a given accuracy, but in their basic implementations they can perform redundant calculations.
+To reduce the number of trials in this class of algorithms, heuristics are often used that reduce the amount of computations \cite{Konnov2025,AGP_ML}.
+Among effective deterministic algorithms for solving global optimization problems, the information-statistical family of global search methods \cite{Strongin2000,Sergeyev2013} can be considered.
+Initially, these algorithms were proposed for solving scalar optimization problems, subsequently, the algorithms were successfully developed to solve multicriterial problems of global search \cite{Konnov2025,MCO_ML_2023}.
+
+In the publication \cite{Konnov2025}, we proposed an approach to solving multicriterial optimization problems that combines a classic global search algorithm with machine learning methods.
+In the framework of the approach, the problem of multicriterial optimization was reduced to solving a series of scalar optimization problems.
+In the process of solving scalar problems , a model based on machine learning methods was built and gradually improved that separated the points of effective and ineffective solutions in the criteria space.
+The constructed model made it possible to reduce the amount of calculations necessary to solve the problem with a given quality.
+Within the framework of this research, we propose the development of the approach for case when the algorithm is used on multi-core computing systems with shared memory.
+
+The article has the following structure.
+Section \ref{sec2} provides a mathematical statement of the problem to be solved.
+Section \ref{sec3} contains a brief description of the developed algorithm and its parallel modification.
+Section \ref{sec4} includes the results of computational experiments demonstrating the effectiveness of the proposed approach.
+
 \section{Problem statement}
 \label{sec2}
+
+The multicriterial optimization (MCO) problem is formulated as follows.
+\begin{equation}
+\label{eq:01}
+  \Phi(y) = (f_1 (y),f_2 (y), \dots, f_s(y))
+\end{equation}
+where $f_i(y)$, $1 \leq i \leq s$, are the criteria of efficiency, $y=(y_1,y_2,\dots,y_N)$ is the vector of varied parameters, N is the dimensionality of the problem solved, and $s$ is the number of varied  criteria.
+It is supposed in this statement that the criteria $f_i(y)$, $1 \leq i \leq s$ are multiextremal functions, given as a ``black box'' and satisfying the Lipschitz condition
+\begin{equation}
+\label{eq:02}
+|f_i (y') - f_i (y'')| \leq L_i \|y' - y''\| ,y',y'' \in D, 1 \leq i \leq s,
+\end{equation}
+where Lipschitz constants $L_i$, $1 \leq i \leq s$, are unknown \textit{a priory}.
+
+Moreover, it is supposed that values of partial criteria $f_i(y)$, $1 \leq i \leq s$, are non-negative and their decreasing corresponds to increasing the solution efficiency.
+If an initial criterion does not satisfy these requirements, it is easily to build a new criterion by means of changing the sign and adding a constant for satisfying the initial requirements.
+Note that the constant can be selected during optimization process \cite{Konnov2025}.
+
+It is supposed that the set of possible values of the optimized parameters $y$ is contained in the $N$-dimensional hypercube $D$:
+\begin{equation}
+\label{eq:03}
+    D=\{y \in R^N : a_i \leq y_i \leq b_i, 1 \leq i \leq N\}
+\end{equation}
+
+As a particular solution to the MCO problem, any effective option can be considered in which it is impossible to reduce the values of all criteria at once $f_i(y)$, $1 \leq i \leq s$, by selection of parameter values $y \in D$.
+In general case, when solving the MCO problem, it may be necessary to find the entire set of Pareto-optimal options \cite{Miettinen1999,Ehrgott2005,Pardalos2017}.
 
@@
 
+In the framework of the research, an algorithm for solving problems of multicriterial optimization with application of machine learning models has been developed.
+The general scheme of the algorithm is presented in Fig.
+\ref{fig1}.
+The algorithm combines elements of a deterministic global search algorithm and machine learning techniques to improve search efficiency.
+The following components can be pointed out in the algorithm.
+\begin{enumerate}
+	\item Block for accumulating and reusing the information about current state of the search.
+	\item Block for building the Pareto set estimation. 
+	\item Block of the global search algorithm. 
+	\item Block of machine learning for building a Pareto set estimation model. 
+	\item Block for the choice of trial points and implementation of computational experiments.
+\end{enumerate}
+
+Numerical solution of optimization problem consists in sequential carrying out trials (computations of criteria values at points of the search domain).
+The first block of the algorithm allows accumulating information about the trials performed and using it to solve problem (\ref{eq:01})-(\ref{eq:03}).
+Block 2 on the base of the accumulated information constructs an approximate solution as an estimate of the Pareto set.
+Block 3 allows efficient selection of trial points based on a global search algorithm.
+Block 4 builds a model for estimating the Pareto set.
+The built model is used to reduce the number of trials required to find the Pareto set with high quality and with fewer trials.
+The fifth block contains an algorithm that allows realizing trials in parallel.
+Let's take a closer look at the computing blocks.
+
 \begin{figure}
 \center
-\includegraphics[width=0.6\textwidth]{fig1.png}
+\includegraphics[width=0.7\textwidth]{fig1.png}
 \caption{Computational scheme of the algorithm for solving multicriterial optimization problems using machine learning models.} \label{fig1}
@@
 \subsection{Block of the global search algorithm}
+\label{subsec31}
+
+The search for a globally optimal solution to problem (\ref{eq:01})-(\ref{eq:03}) consists of the following basic steps:
+\begin{enumerate}
+	\item scalarization of efficiency criteria;
+	\item dimensionality reduction;
+	\item optimization on the base of the global search algorithm.
+\end{enumerate}
+
+\textbf{Scalarization of efficiency criteria.} It is possible to find the Pareto set either solving the problem in the original formulation \cite{Evtushenko2014,DPA02,Durillo2010,Mostaghim2007,Nebro2009,RC05,Zitzler2001}, or reducing the original problem to a series of scalar problems to find effective solutions \cite{Pardalos2017,Konnov2025,Gergel2019_2,Gergel2018,GergelKozinov2020}.
+
+In the framework of the developed approach \cite{Konnov2025}, the second option with scalarization of criteria is used. To build a good estimate of the Pareto set, a family of problems
 
 \subsection{Block for accumulating and reusing the information about current state of the search}
+\label{subsec32}
 
 \subsection{Block for building the Pareto set estimation}
+\label{subsec33}
 
@@
 \subsection{Block of  building the machine learning for estimation of the Pareto set}
+\label{subsec34}
 
@@
 \subsection{Block for the choice of trial points and implementation of computational experiments}
+\label{subsec35}
 

### 2025-04-06 2025/RSCD_MCO/RSD-2025.tex
MCO весь текст 15 страниц

@@
 
-\cite{Miettinen1999,Ehrgott2005,Pardalos2017,Konnov2025,Evtushenko2014,DPA02,Durillo2010,Mostaghim2007,Nebro2009,RC05,Zitzler2001,Gergel2019_2,Gergel2018,GergelKozinov2020,Marler2004,Strongin2000,Sergeyev2013,MCO_ML_2023,scikit-learn,PyTorch,ioptmco,HV,pymoo,AGP_ML}
+%\cite{Miettinen1999,Ehrgott2005,Pardalos2017,Konnov2025,Evtushenko2014,DPA02,Durillo2010,Mostaghim2007,Nebro2009,RC05,Zitzler2001,Gergel2019_2,Gergel2018,GergelKozinov2020,Marler2004,Strongin2000,Sergeyev2013,MCO_ML_2023,scikit-learn,PyTorch,ioptmco,HV,pymoo,AGP_ML}
 
@@
 In the framework of the developed approach \cite{Konnov2025}, the second option with scalarization of criteria is used. To build a good estimate of the Pareto set, a family of problems
+\begin{equation}
+\label{eq:04}
+\min {\varphi_j (y)}=F(\lambda_j,y),F(\lambda_j,y) \in \{F(\lambda_1,y),F(\lambda_2,y),\dots,F(\lambda_\tau,y)\},y\in D,
+\end{equation}
+is solved,  where $F(\lambda_j,y)$, $1 \leq j \leq \tau$, is a multiextremal function, obtained on the base of scalarization of criteria $f_i(y)$, $1 \leq i \leq s$.
+Here $\lambda_j=\{\lambda_{i,j}\}$, $1 \leq i \leq s$, $1 \leq j \leq \tau$, are significance indicators for each criterion, s is the number of criteria, $\tau$ is the number of scalar optimization problems.
+There are different ways to build functions $F(\lambda_j,y)$ \cite{Miettinen1999,Konnov2025,Marler2004}.
+The performed computational experiments used minimax convolution of criteria
+\begin{equation}
+\label{eq:05}
+F(\lambda_j, y) = \max_{1 \leq i \leq s} \left(\lambda_{i,j} f_i (y)\right), y \in D, \sum_{i=1}^s {\lambda_{i,j}} = 1, 0 \leq \lambda_{i,j} \leq 1.
+\end{equation}
+
+\textbf{Dimensionality reduction.} For solving optimization problems (\ref{eq:04}) the dimensionality reduction on the base of Peano curves $y(x)$ that map unambiguously and continuously the segment $[0,1]$ to the $N$-dimensional hypercube $D$ is used \cite{Konnov2025,Gergel2019_2,GergelKozinov2020}.
+As a result of this reduction, multidimensional global optimization problems (\ref{eq:04}) are reduced to one-dimensional subproblems
+\begin{equation}
+\label{eq:06}
+\min_x {\varphi_j(y(x))} = \min_y {\varphi_j(y)}, x \in [0,1], y \in D, 1 \leq j \leq \tau
+\end{equation}
+
+Note that the reduced one-dimensional functions $\varphi_j (y(x))$ from (\ref{eq:06}) satisfy the H{\" o}lder condition, i.e.
+\begin{equation}
+\label{eq:07}
+|\varphi_j(y(x')) - \varphi_j(y(x''))| \leq H |x' - x''|^{1/N} , x \in [0,1],
+\end{equation}
+where the constant $H$ is given as $H=2L\sqrt{N+3}$, and $L$ is the Lipschitz constant from (\ref{eq:02}).
+
+\textbf{Global search algorithm.} Information-statistical algorithm of global search (shortly GSA) \cite{Konnov2025,Gergel2019_2,Gergel2018,GergelKozinov2020,Strongin2000,Sergeyev2013} is used to solve problems (\ref{eq:06}).
+The computational scheme of GSA is as follows.
+At each iteration of the global search, one trial (calculation of the value of the optimized function $\varphi_j (x)$ from (\ref{eq:06})) is executed.
+In the first iteration, two tests are carried out at the ends of the interval $x^0=0$,$x^1=1$.
+Let $k$, $k>2$, iterations of the global search be performed.
+The selection of the next test point is based on the following rules.
+
+\textit{Rule 1.} Renumber search iteration points with subscripts in order of increasing coordinate values
+\begin{equation}
+    \label{eq:08}
+    0 = x_0 < x_1 < \dots < x_i < \dots < x_{k} = 1
+\end{equation}
+\textit{Rule 2.} Calculate the current estimate of the H{\" o}lder constant from (\ref{eq:07}) for the reduced function $\varphi_j (x)$:
+\begin{equation}
+    \label{eq:9}
+		\begin{matrix}
+		m=\begin{cases}
+				\begin{matrix}
+					 r M, & M >0 \\
+					 1, & M = 0 
+				\end{matrix} \; , 
+			\end{cases}
+		M = \max_{1 \leq i \leq k} \frac{| z_i - z_{i-1}|}{\varrho_i}, \\
+		z_i = \varphi_j( y(x_i) ), \varrho_i=\sqrt[N]{x_i-x_{i-1}}
+		\end{matrix}
+\end{equation}
+where $r$, $r>1$, is the reliability parameter of the algorithm.
+
+\textit{Rule 3.} Compute characteristics $R(i)$ for each interval $(x_{i-1}, x_i)$, $1\leq i \leq k$, according to the expression
+\begin{equation}
+    \label{eq:10}
+    R(i) = \varrho_i + \frac{(z_i-z_{i-1})^2}{m^2 \varrho_i} - \frac{2 (z_i+z_{i-1})}{m}, 1 \leq i \leq k.
+\end{equation}
+
+\textit{Rule 4.} Select the interval with the maximal characteristic $R(i)$ from (\ref{eq:10})
+\begin{equation}
+    \label{eq:11}
+    R(t) = arg\max_{1 \leq i \leq k} {R(i)}.
+\end{equation}
+
+\textit{Rule 5.} Execute the next trial in the interval with the maximal characteristic value, according to the expression
+\begin{equation}
+    \label{eq:12}
+    x^{k+1} = \frac{x_t + x_{t-1}}{2} - sign(z_t - z_{t-1}) \frac{1}{2r} \left[\frac{|z_t - z_{t-1}|}{m} \right]^N
+\end{equation}
+
+The stopping condition of the algorithm, according to which the execution of the algorithm for the current global search problem (\ref{eq:06}) is terminated, consists in achieving the required accuracy of solving the problem, i.e.
+\begin{equation}
+    \label{eq:13}
+    \varrho_t < \varepsilon.
+\end{equation}
 
@@
 
+As mentioned earlier, the numerical solution to optimization problems consists in sequential realization of trials $z^i=\varphi_j (y^i)$ from (\ref{eq:04}) at points $y^i$, $1 \leq i \leq k$, of the search domain $D$.
+The obtained data can be presented in the form of \textit{the search information set} (SIS):
+\begin{equation}
+    \label{eq:14}
+    \Omega_k=\{\mu^i=(y^i,f^i=f(y^i))^T: 0 \leq i \leq k\}
+\end{equation}
+
+SIS contains all available information about the solved problem MCO (\ref{eq:01})-(\ref{eq:03}).
+Improved search efficiency can be achieved on the base of information available in $\Omega_k$.
+In our approach SIS is used:
+\begin{enumerate}
+	\item	when the scalar statement of the problem (\ref{eq:04}) is changed,
+	\item	during building a Pareto set estimate, 
+	\item	in the process of building a machine learning model based on the Pareto set.
+\end{enumerate}
+
+Within item 1 the following sequence of actions is performed.
+When task (\ref{eq:06}) is changed, new coefficients $\lambda_{j+1}$ from (\ref{eq:04}) are selected.
+Further, scalarization of the vector criterion (\ref{eq:04}) and dimensionality reduction (\ref{eq:06}) transform SIS from (\ref{eq:14}) to the form used in the optimization process (\ref{eq:08})-(\ref{eq:13}) for the current problem (\ref{eq:04}).
+Transformation uses already calculated criteria values and new coefficients $\lambda_{j+1}$.
+As a consequence, the recount does not take long time.
+Thus, information about the calculated criteria is not lost when changing the task statement, but is reused.
+This approach has shown its effectiveness in solving many problems \cite{Konnov2025,Gergel2018,GergelKozinov2020}.
+
+The calculations fulfilled under items 2 and 3 are described in sections \ref{subsec33} and \ref{subsec34} respectively.
+
 \subsection{Block for building the Pareto set estimation}
 \label{subsec33}
+
+Accumulation of search information allows more accurate estimation of the Pareto set.
+In solving the series of problems (\ref{eq:04}), after each trial performed or task setting change, the current Pareto set is selected on the base of all the trials executed.
+Thus, even when solving one optimization problem from (\ref{eq:04}), the Pareto set evaluation contains several variants of solutions, from which the researcher can choose the best variant for himself (see Fig.
+\ref{fig2}, blue dots).
+Solving additional tasks (\ref{eq:04}) makes it possible to improve the found solution (see Fig.
+\ref{fig2}).
 
@@
 
+To improve the search efficiency within the developed scheme, the calculation of characteristics for intervals (\ref{eq:10}) is changed to the following dependence \cite{Konnov2025}:
+\begin{equation}
+    \label{eq:15}
+    R(i) = R_{GSA} (i) +  \alpha R_{PS} (i)
+\end{equation}
+
+The characteristic $R_{GSA}(i)$ from (\ref{eq:15}) is used to find the global minimum in the current optimization problem.
+The method of calculating $R_{GSA}(i)$ corresponds to rule 3 of the GSA in formula (\ref{eq:10}).
+The value of $R_{PS}(i)$ is calculated on the base of the found Pareto set in the current optimization step and links solvable optimization problems from the set (\ref{eq:04}).
+The coefficient $\alpha$ changes the degree of influence of the characteristic  $R_{PS}(i)$ on the choice of the next point.
+At low $\alpha$ values, the algorithm executes more trials to find the optimal solution of the current optimization problem (\ref{eq:04}).
+At high $\alpha$ values, a new trial is carried out in the vicinity of the Pareto set.
+
+To calculate the value of $R_{PS}(i)$, we propose using machine learning methods.
+Various algorithms for calculation of  $R_{PS}(i)$ have been proposed in \cite{Konnov2025}.
+All proposed methods use the following scheme.
+In the course of solving the MCO problem, after every changing the coefficients  $\lambda_j$, $1 \leq j \leq \tau$, and therefore, moving to a new optimization problem from (\ref{eq:06}), all obtained trial points (\ref{eq:14}) are divided into two classes.
+The first class includes points belonging to the Pareto set.
+All other trial points belong to the second class.
+The quantity of points that do not belong to the Pareto set, as a rule, is significantly larger; therefore, the classes are not balanced.
+For more efficient work of machine learning algorithms, weights corresponding to the significance of trial points are set for each class, if possible.
+After training the model in the selected classes, the following approaches are used for building $R_{PS}(i)$.
+\begin{enumerate}
+	\item When using the Support Vector Classification (SVC) model with a linear core, $R_{PS}(i)$ is considered on the base of the distance to the dividing hyperplane.
+	\item	When using the SVC model with an arbitrary core, $R_{PS}(i)$ RP is built according to probabilities of belonging trial points to the first class (Pareto set).
+	\item Neural network models are used as a new way to evaluate $R_{PS}(i)$. The method of calculating $R_{PS}(i)$ is similar to step 2.
+\end{enumerate}
+
+In the framework of the first approach, an SVC model with a linear core is built for computation $R_{PS}(i)$.
+Information is extracted from the model to construct a dividing hyperplane equation between the selected point classes in the criteria value space.
+The characteristic $R_{PS}(i)$ of the interval $[x_{i-1},x_i ]$ is considered as the sum of the distances of the criteria values  from the points $x_{i-1}$, $x_i$ to the hyperplane.
+The value below the hyperplane is taken with a plus sign, and above the hyperplane  with a minus sign.
+
+In the second approach, the probability of belonging  each of the points to the Pareto region is used instead of distances.
+$R_{PS}(i)$ is considered as the arithmetic mean of the calculated probability values for the points $x_{i-1}$ and $x_i$.
+This approach is more universal, since the counting method does not depend on the selected kernel.
+The method of calculating $R_{PS}(i)$ is described more detailed in \cite{Konnov2025}.
+
+The third approach is based on the second one and uses new models based on fully connected neural networks. Two variants of models were considered.
+\begin{enumerate}
+	\item Variant with two hidden layers (Fig.
+\ref{fig3}, a).
+In this case, the Input layer consists of $N$ nodes, where the number $N$ must be equal to the number of criteria in the problem being solved.
+The first hidden layer contains 128 nodes, the second hidden layer includes 64 nodes, output layer consists of 2 nodes, each of which denotes the probability of assigning a point to one of the classes describing belonging to the Pareto set.
+The ReLU function was used as the activation function after the Input layer and the first hidden layer.
+To obtain the probability before the output layer, the softmax function is applied to the data.
+	\item Variant with one hidden layer (Fig.
+\ref{fig3}, b).
+This is a simpler option, in which the Input layer also consists of N nodes, where N is equal to the number of criteria.
+The next (hidden) layer contains 100 nodes.
+The last (output) layer also consists of 2 layers.
+Similar to the previous case, the activation function is ReLU, and before the output layer, the softmax function is applied.
+\end{enumerate}
+
 \begin{figure}
@@
 \end{figure}
+
+Figure \ref{fig4} shows an example of built functions $R_{PS}(i)$ in the criteria domain.
+Red denotes the points assigned by the algorithm to the Pareto set.
+Graph (a) refers to the first approach with the calculation of distances to the hyperplane.
+Plots (b), (c), (d) demonstrate results for the second approach using SVC models to calculate probabilities.
+Graphs (e), (f) show heat maps for the third approach using neural network models.
+
+From the figure, it can be seen that the function $R_{PS}(i)$ for each of the considered variants differs in the quality of the evaluation of Pareto sets.
+When choosing a machine learning model, it should be taken into account that building a model with a good description of the domain may require more computing resources.
+An analysis of the impact of the selected model on the quality of the resulting solution is presented in Section \ref{sec4} with the results of computational experiments.
+Note that numerous computational experiments have shown that changing the method of calculating characteristic (\ref{eq:15}) leads to a significant increase in the efficiency of the algorithm as a whole \cite{Konnov2025,MCO_ML_2023}.
 
@@
 
+To further improve the search efficiency, a parallel calculation scheme has been implemented as a new element of the algorithm for solving multicriterial optimization problems using machine learning \cite{ioptmco}.
+
+In global search algorithms, the traditional method of parallelization by dividing the search area into parts is not effective, since the minimum values of the criteria can fall into a limited number of subregions, therefore, many trials would not improve the Pareto set estimation.
+In this work, the classic master-working circuit is used as a parallel scheme.
+
+Workflows are waiting for a new trial point or command to complete calculations. When a point is received, criteria calculations are performed and the master is notified of the results.
+
+The scheme of the master is as follows.
+Let there be $p$ workflows.
+The first $p$ trials are executed in parallel at points $x_0=0<x_1<\dots<x_i<\dots<x_{p-1}=1$, $0<i<p$, uniformly distributed on the segment $[0,1]$.
+Further, in the case when all workflows are busy, the master is in the waiting state.
+When information from the workflow appears, the master obtains the value of the criteria from (\ref{eq:06}) and updates the set of search information (\ref{eq:14}), calculates the estimate of the H{\" o}lder constant (\ref{eq:07}) as well as the values of the characteristics of the intervals (\ref{eq:15}).
+As in the case of a sequential implementation, it selects the interval with the maximal characteristic, and also computes the trial point (\ref{eq:12}).
+The obtained value is sent to the workflow to continue the computations.
+To avoid repeated calculations, intervals already involved in calculations on workflows are not involved in the calculation of characteristics.
+
+As in the sequential algorithm, stopping and evaluating the resulting solution is carried out according to rule (\ref{eq:13}).
+A detailed analysis of the effectiveness of the given parallel implementation is presented in Section \ref{sec4} with computational experiments.
+
 \section{Results of computational experiments}
 \label{sec4}
+
+Computational experiments were carried out on the nodes of the Lobachevsky supercomputer of Nizhny Novgorod State University with two processors Intel Xeon Silver 4310T 2.3 GHz, 64 GB RAM.
+Machine learning libraries SciKit-Learn 0.24.2 \cite{scikit-learn} and PyTorch 2.5.1 \cite{PyTorch} were used in experiments.
+The algorithm proposed in the work was implemented in Python using the framework iOpt \cite{ioptmco}.
+
+As a criterion for the quality of the found Pareto set, the indicator HW \cite{Evtushenko2014,pymoo,AGP_ML} was applied.
+Higher value of the indicator corresponds to better solution.
+To assess the performance of the algorithm, the number of trials carried out when solving the problem are given.
+However, when solving complex optimization problems, one trial can take a significant time.
+In such conditions, the reduction of the number of trials reduces the total computation time in general.
+In the presented experimental results, the number of trials spent on solving each problem and the HW indicators were averaged according to the number of solved problems.
+
+To estimate the quality of the resulting solutions, a number of computational experiments were performed to solve 100 unconstrained two-criteria two-dimensional MCO problems.
+The following multi-extremal functions \cite{Gergel2019_2}:
+\begin{equation}
+    \label{eq:16}
+		\begin{matrix}
+		  f(y)= -(AB + CD)^{1/2}, \\
+			AB =(\sum_{i=1}^7{\sum_{j=1}^7{[A_{ij} a_{ij} (y_1,y_2) + B_{ij} b_{ij} (y_1,y_2)])}})^2, \\
+			CD =(\sum_{i=1}^7{\sum_{j=1}^7{[C_{ij} a_{ij} (y_1,y_2) - D_{ij} b_{ij} (y_1,y_2)])}})^2, \\
+			a_{ij} (y_1,y_2) = \sin(\pi i y_1) \sin(\pi j y_2), \\
+			b_{ij} (y_1,y_2) = \cos(\pi i y_1) \cos(\pi j y_2),
+		\end{matrix}
+\end{equation}
+were used as partial criteria in the two-criteria statement, where $0 \leq y_1, y_2 \leq 1$, parameters $-1 \leq A_{ij},B_{ij},C_{ij},D_{ij} \leq 1$ are independent uniformly distributed random values.
+
+Results of the experiments for the following algorithms are presented (see subsection \ref{subsec34}):
+\begin{enumerate}
+	\item MGSA -- machine learning elements are not used,
+	\item MGSA Dist -- the modified algorithm with computation $R_{PS}(i)$ from (\ref{eq:15}) based on distances,
+	\item MGSA Linear SVC, MGSA Poly SVC, MGSA Rbf SVC -- the algorithm modification with computation $R_{PS}(i)$ from (\ref{eq:15}) based on probabilities with different SVC algorithm core,
+	\item MGSA NN -- the algorithm modification with computation $R_{PS}(i)$ from (\ref{eq:15}) on the base of probabilities obtained using neural networks.
+\end{enumerate}
+
+Fig.
+\ref{fig5} shows the dependence of the average value of the HW index on the average number of trials for the sequential algorithm.
+Several solution accuracy values $\varepsilon = \{0.1,0.05,0.01\}$ were fixed for plotting the graph for algorithm (\ref{eq:13}).
+100 problems were solved for each accuracy value and the average results were plotted on a graph.
+The position of the resulting curve for the chosen algorithm to the left and above corresponds to the best behavior of the algorithm as a whole.
+During the experiments the parameter $\alpha$ from (15) varied, Fig.
+\ref{fig5} shows only the best results for each of the variants.
 
@@
 \end{figure}
+
+Based on the results of experiments with sequential implementation, all approaches using machine learning models exceed the MGSA algorithm by more than two times.
+Modifications of the algorithm can be distinguished that demonstrate the best results both in terms of the average number of iterations and in terms of the average HW.
+These include approaches using SVC (Linear, Rbf cores), a distance calculation approach, and neural networks with two hidden layers.
+Modification with one hidden layer shows slightly worse results.
+Modification  MGSA Poly SVC with polynomial core showed least improvement over MGSA.
+It is also worth noting that the modification of the MGSA Rbf SVC algorithm in the case of low accuracy demonstrates the best quality.
+
+Next, consider a parallel version of the algorithm (see subsection \ref{subsec35}).
+Table \ref{tab:1} shows the results of the average value of the trial number per process by tasks for each of the models in dependence on the number of processes.
+Accuracy $\varepsilon=0.01$ was used in the experiment.
+The number of processes used is indicated in the first row of the table.
+The three best modifications of the algorithm in terms of the average number of trials per process are shown in bold.
 
@@
 \textbf{Number of processes}            & \textbf{1}      & \textbf{2}      & \textbf{4}      & \textbf{8}      & \textbf{16}    & \textbf{20}    \\ \hline
-MGSA                                    & 2269.82         & 1385.02         & 746.57          & 373.36          & 202            & 143.52         \\
-\textbf{MGSA   Linear SVC,  a = 0.05}   & \textbf{686.13} & \textbf{408.79} & \textbf{213.6}  & \textbf{111.85} & \textbf{70.95} & \textbf{62.16} \\
-MGSA   Poly SVC, a = 0.05               & 909.59          & 544.19          & 278.62          & 153.64          & 95.37          & 77.94          \\
-MGSA   RBF SVC, a = 0.04                & 652.74          & 446.84          & 233.81          & 119.84          & 72.36          & 62.37          \\
-\textbf{MGSA   Dist, a = 0.01}          & \textbf{659.98} & \textbf{402.9}  & \textbf{211.33} & \textbf{112.38} & \textbf{70.79} & \textbf{62.00} \\
-MGSA   NN 1 layer, a = 0.05             & 795.16          & 446.25          & 232.08          & 117.16          & 75.19          & 62.70          \\
-\textbf{MGSA   NN 2 layers, a =   0.03} & \textbf{682.12} & \textbf{401.74} & \textbf{219.62} & \textbf{113.32} & \textbf{73.51} & \textbf{61.32} \\ \hline
+MGSA                                    & 2269.82         & 1385.02         & 746.57          & 373.36          & 202.01            & 143.52         \\
+\textbf{MGSA   Linear SVC,  $\alpha$ = 0.05}   & \textbf{686.13} & \textbf{408.79} & \textbf{213.6}  & \textbf{111.85} & \textbf{70.95} & \textbf{62.16} \\
+MGSA Poly SVC, $\alpha$ = 0.05               & 909.59          & 544.19          & 278.62          & 153.64          & 95.37          & 77.94          \\
+MGSA RBF SVC, $\alpha$ = 0.04                & 652.74          & 446.84          & 233.81          & 119.84          & 72.36          & 62.37          \\
+\textbf{MGSA Dist, $\alpha$ = 0.01}          & \textbf{659.98} & \textbf{402.9}  & \textbf{211.33} & \textbf{112.38} & \textbf{70.79} & \textbf{62.00} \\
+MGSA NN 1 layer, $\alpha$ = 0.05             & 795.16          & 446.25          & 232.08          & 117.16          & 75.19          & 62.70          \\
+\textbf{MGSA NN 2 layers, $\alpha$ =   0.03} & \textbf{682.12} & \textbf{401.74} & \textbf{219.62} & \textbf{113.32} & \textbf{73.51} & \textbf{61.32} \\ \hline
 \end{tabular}%
@@
 \end{table}
+
+According to Table \ref{tab:1}, MGSA Dist, MGSA Linear SVC and MGSA NN algorithms with two hidden layers demonstrate the best results.
+To analyze the quality of the resulting solution, the values   of the HW index are given in Table \ref{tab:2}.
+The table also shows three best modifications.
 
@@
 \textbf{MGSA}                         & \textbf{114.91} & \textbf{114.79} & \textbf{114.84} & \textbf{114.90} & \textbf{115.00} & \textbf{114.96} \\
-MGSA   Linear SVC, a = 0.05           & 114.44          & 114.65          & 114.72          & 114.74          & 114.91          & 114.96          \\
-\textbf{MGSA   Poly SVC,    a = 0.05} & \textbf{114.61} & \textbf{114.64} & \textbf{114.71} & \textbf{114.8}  & \textbf{114.95} & \textbf{115.01} \\
-MGSA   RBF SVC,    a = 0.04           & 114.54          & 114.54          & 114.73          & 114.89          & 114.92          & 114.96          \\
-MGSA   Dist, a = 0.01                 & 114.66          & 114.55          & 114.69          & 114.73          & 114.92          & 114.93          \\
-MGSA   NN 1 layer, a = 0.05           & 114.77          & 114.61          & 114.72          & 114.81          & 114.96          & 114.98          \\
-\textbf{MGSA   NN 2 layers, a = 0.03} & \textbf{114.74} & \textbf{114.54} & \textbf{114.71} & \textbf{114.76} & \textbf{114.98} & \textbf{115.01} \\ \hline
+MGSA Linear SVC, $\alpha$ = 0.05           & 114.44          & 114.65          & 114.72          & 114.74          & 114.91          & 114.96          \\
+\textbf{MGSA   Poly SVC, $\alpha$ = 0.05} & \textbf{114.61} & \textbf{114.64} & \textbf{114.71} & \textbf{114.8}  & \textbf{114.95} & \textbf{115.01} \\
+MGSA RBF SVC, $\alpha$ = 0.04           & 114.54          & 114.54          & 114.73          & 114.89          & 114.92          & 114.96          \\
+MGSA Dist, $\alpha$ = 0.01                 & 114.66          & 114.55          & 114.69          & 114.73          & 114.92          & 114.93          \\
+MGSA NN 1 layer, $\alpha$ = 0.05           & 114.77          & 114.61          & 114.72          & 114.81          & 114.96          & 114.98          \\
+\textbf{MGSA NN 2 layers, $\alpha$ = 0.03} & \textbf{114.74} & \textbf{114.54} & \textbf{114.71} & \textbf{114.76} & \textbf{114.98} & \textbf{115.01} \\ \hline
 \end{tabular}%
@@
 \end{table}
+
+From Table \ref{tab:2} as a whole, it can be concluded that parallel modification of the algorithm does not worsen the quality of the found solution.
+For a more detailed productivity analysis let's plot the graph by analogy with Fig.
+\ref{fig5} for the case of 20 processes.
 
@@
 \end{figure}
+
+Fig.
+\ref{fig6} demonstrates that as the number of processes used increases, all algorithms are significantly ahead of the MGSA method, which does not use machine learning models, in both the average number of trials per process and the HW indicator, especially if high solution accuracy is to be ensured.
+All modifications of the algorithm except MGSA Poly SVC and MGSA NN with one layer show similar quality.
+A simpler modification MGSA NN with one hidden layer with a small number of trials lags behind in quality and aligns with the rest when increasing the number of trials.
+This may indicate that this model can be useful when high-precision calculations are needed, in other cases it is recommended to use other machine learning models.
+At the same time, a simple model can learn much faster.
+
+Fig.
+\ref{fig7} shows the value of the HW index depending on the number of trials when the number of used processes changes.
+Since MGSA NN 2 layers demonstrates better results than MGSA NN 1 layer, the information about the last modification are not given.
+According to Fig.
+\ref{fig7}, all presented models show a significant increase in the HW index and a decrease in the trial number per process when increasing the number of processes.
+Approach without using machine learning methods demonstrate similar dynamics when the target accuracy is high, however, at lower values, changes in the number of counted criteria on one process become less noticeable.
+It is also worth noting that, unlike the basic MGSA algorithm, all modifications with machine learning elements do not lose quality when switching to a parallel implementation.
 

### 2025-04-06 2025/RSCD_MCO/RSD-2025.tex
fix MCO HW->HV

@@
 
-As a criterion for the quality of the found Pareto set, the indicator HW \cite{Evtushenko2014,pymoo,AGP_ML} was applied.
+As a criterion for the quality of the found Pareto set, the indicator HV \cite{Evtushenko2014,pymoo,AGP_ML} was applied.
 Higher value of the indicator corresponds to better solution.
@@
 In such conditions, the reduction of the number of trials reduces the total computation time in general.
-In the presented experimental results, the number of trials spent on solving each problem and the HW indicators were averaged according to the number of solved problems.
+In the presented experimental results, the number of trials spent on solving each problem and the HV indicators were averaged according to the number of solved problems.
 
@@
 Fig.
-\ref{fig5} shows the dependence of the average value of the HW index on the average number of trials for the sequential algorithm.
+\ref{fig5} shows the dependence of the average value of the HV index on the average number of trials for the sequential algorithm.
 Several solution accuracy values $\varepsilon = \{0.1,0.05,0.01\}$ were fixed for plotting the graph for algorithm (\ref{eq:13}).
@@
 \includegraphics[width=0.8\textwidth]{fig5.png}
-\caption{Comparison of HW index values for sequential algorithm implementation.} \label{fig5}
+\caption{Comparison of HV index values for sequential algorithm implementation.} \label{fig5}
 \end{figure}
@@
 Based on the results of experiments with sequential implementation, all approaches using machine learning models exceed the MGSA algorithm by more than two times.
-Modifications of the algorithm can be distinguished that demonstrate the best results both in terms of the average number of iterations and in terms of the average HW.
+Modifications of the algorithm can be distinguished that demonstrate the best results both in terms of the average number of iterations and in terms of the average HV.
 These include approaches using SVC (Linear, Rbf cores), a distance calculation approach, and neural networks with two hidden layers.
@@
 According to Table \ref{tab:1}, MGSA Dist, MGSA Linear SVC and MGSA NN algorithms with two hidden layers demonstrate the best results.
-To analyze the quality of the resulting solution, the values   of the HW index are given in Table \ref{tab:2}.
+To analyze the quality of the resulting solution, the values   of the HV index are given in Table \ref{tab:2}.
 The table also shows three best modifications.
@@
 \centering
-\caption{Average HW Index in dependence on number of processes}
+\caption{Average HV Index in dependence on number of processes}
 \label{tab:2}
@@
 \includegraphics[width=0.8\textwidth]{fig6.png}
-\caption{Dependence of the mean value of the index HW from the trial number per process for the case of 20 processes.} \label{fig6}
+\caption{Dependence of the mean value of the index HV from the trial number per process for the case of 20 processes.} \label{fig6}
 \end{figure}
@@
 Fig.
-\ref{fig6} demonstrates that as the number of processes used increases, all algorithms are significantly ahead of the MGSA method, which does not use machine learning models, in both the average number of trials per process and the HW indicator, especially if high solution accuracy is to be ensured.
+\ref{fig6} demonstrates that as the number of processes used increases, all algorithms are significantly ahead of the MGSA method, which does not use machine learning models, in both the average number of trials per process and the HV indicator, especially if high solution accuracy is to be ensured.
 All modifications of the algorithm except MGSA Poly SVC and MGSA NN with one layer show similar quality.
@@
 Fig.
-\ref{fig7} shows the value of the HW index depending on the number of trials when the number of used processes changes.
+\ref{fig7} shows the value of the HV index depending on the number of trials when the number of used processes changes.
 Since MGSA NN 2 layers demonstrates better results than MGSA NN 1 layer, the information about the last modification are not given.
+
 According to Fig.
-\ref{fig7}, all presented models show a significant increase in the HW index and a decrease in the trial number per process when increasing the number of processes.
+\ref{fig7}, all presented models show a significant increase in the HV index and a decrease in the trial number per process when increasing the number of processes.
 Approach without using machine learning methods demonstrate similar dynamics when the target accuracy is high, however, at lower values, changes in the number of counted criteria on one process become less noticeable.
@@
 \includegraphics[width=0.8\textwidth]{fig7.png}
-\caption{Dependence of the HW index on the number of trials executed for different number of processes.} \label{fig7}
+\caption{Dependence of the HV index on the number of trials executed for different number of processes.} \label{fig7}
 \end{figure}

### 2025-04-09 2025/RuSCDays/RSD-2025.tex
add references and correct typos

@@
 Для демонстрации эффективности распараллеливания проведены вычислительные эксперименты в задаче настройки гиперпараметров методов искусственного интеллекта, используемых для предсказания значений временного ряда.
-The paper considers a parallel algorithm for solving global optimization problems with a partially defined objective function of the “black box” type.
-Such problems arise in the process of selecting optimal values of hyperpa-rameters used in machine learning and artificial intelligence methods.
-The al-gorithm employs a dimensionality reduction scheme based on Peano curves in combination with an asynchronous scheme of computation parallelization.
-To demonstrate the effectiveness of parallelization, computational experi-ments were conducted for the task of tuning the hyperparameters of artificial intelligence methods used to predict the values of a time series.
+The paper considers a parallel algorithm for solving global optimization problems with a partially defined objective function of the ``black box'' type.
+Such problems arise in the process of selecting optimal values of hyperparameters used in machine learning and artificial intelligence methods.
+The algorithm employs a dimensionality reduction scheme based on Peano curves in combination with an asynchronous scheme of parallel computing.
+To demonstrate the effectiveness of parallelization, computational experiments were conducted for the task of tuning the hyperparameters of artificial intelligence methods used to predict the values of a time series.
 
@@
 Currently, machine learning and artificial intelligence methods are used in many academic and applied fields.
-They are successfully applied to address problems that until recently were solved only by classical numerical methods [REFS].
+They are successfully applied to address problems that until recently were solved only by classical numerical methods \cite{Blechschmidt2021,Seleznev2019,Xu2020}.
 
@@
 Most artificial intelligence methods involve hyperparameters, and the choice of their specific values may determine their performance.
-For example, the quality of object classification using the support vector method, which is traditionally measured by the metric $F_1$ [REF], is significantly affected by two parameters of the method: the regularization coefficient and the kernel coefficient.
+For example, the quality of object classification using the support vector machine, which is traditionally measured by the $F_1$-score \cite{NIPS2015}, is significantly affected by two parameters of the method: the regularization coefficient and the kernel coefficient.
 However, the default values geared to the average case will not be the best for a particular class of problems to be solved.
@@
 Thus, there arises the hyperparameter optimization (HPO) problem, which from the mathematical point of view is a global optimization problem.
-In this case, it is typical to have not only continuous but also categorical parameters (parameters having a pre-fixed set of values), i.e.
+In this case, it is typical to have not only continuous but also categorical parameters (parameters having a prefixed set of values), i.e.
 there is a need to use global optimization methods capable of solving problems with mixed parameters.
@@
 The traditional complexity of HPO problems is due to the considerable time required to evaluate a criterion for a particular combination of parameters.
-Therefore, metaheuristic algorithms, which require a large number of search tri-als, can find only a limited use for such problems.
-A more promising approach is to use deterministic Lipschitz optimization algorithms [Evtushenko, Sergeev, Strongin], which are superior to heuristic methods in terms of performance quality [Sergeev].
-For Lipschitz global optimization methods, a number of paral-lelization approaches have been proposed to ensure efficient operation in super-computer systems [Evtushenko, Ziliskas, Gergel].
+Therefore, metaheuristic algorithms, which require a large number of search trials, can find only a limited use for such problems.
+A more promising approach is to use deterministic Lipschitz optimization algorithms \cite{Evtushenko2013,Sergeyev2017,Strongin2000}, which are superior to heuristic methods in terms of performance quality \cite{Sergeyev2018}.
+For Lipschitz global optimization methods, a number of parallelization approaches have been proposed to ensure efficient operation in supercomputer systems \cite{Evtushenko2009,Gegrel2021,Paulavicius2011}.
 
@@
 область допустимых сочетаний параметров является заранее неопределенной.
-One more difficulty of the class of HPO problems being considered is the pres-ence of such combinations of hyperparameter values for which the method under study does not work in principle in some (not known in advance) subdomains of the search domain.
+One more difficulty of the class of HPO problems being considered is the presence of such combinations of hyperparameter values for which the method under study does not work in principle in some (not known in advance) subdomains of the search domain.
 With such values of parameters it is impossible to perform the calculation correctly and calculate the value of the efficiency criterion.
@@
 This paper continues to develop the approach to solving Lipschitz global optimization problems in parallel as proposed in [REFS ON OUR WORKS].
-The new element is the combination of parallelization methods and the scheme for handling par-tially computable functions proposed in [Usova].
-The main part of the publica-tion has the following structure.
-Section 2 presents the mathematical formulation of the problem to be solved.
-Section 3 contains a description of the parallel global search algorithm.
-Section 4 describes the method for solving problems with mixed (discrete and continuous) parameters, and also discusses the method for handling functions that are not defined in the entire search domain.
-Section 5 gives the results of computational experiments demonstrating the effectiveness of the proposed approach in solving the problem of optimizing the hyperparameters of time series forecasting methods.
+The new element is the combination of parallelization methods and the scheme for handling partially computable functions proposed in \cite{Usova2024}.
+The main part of the publication has the following structure.
+Section \ref{sec:PS} presents the mathematical formulation of the problem to be solved.
+Section \ref{sec:PA} contains a description of the parallel global search algorithm.
+Section \ref{sec:MP} describes the method for solving problems with mixed (discrete and continuous) parameters, and also discusses the method for handling functions that are not defined in the entire search domain.
+Section \ref{sec:RCE} gives the results of computational experiments demonstrating the effectiveness of the proposed approach in solving the problem of optimizing the hyperparameters of time series forecasting methods.
 
-\section{Problem Statement}
+\section{Problem Statement}\label{sec:PS}
 %(рус.) В начале рассмотрим задачу оптимизации в простой постановке, только с непрерывными параметрами. Пусть требуется найти минимум функции $\phi(y)$, определенной в области $D$
@@
 A well-known way to solve Lipschitz global optimization problems is to reduce them (explicitly or implicitly) to the solution of one-dimensional problems.
-Such a reduction can be performed, for example, by using diagonal or simplex partitions of the search domain [REFS].
+Such a reduction can be performed, for example, by using diagonal or simplicial partitions of the search domain \cite{PaulaviciusZilinskas2014,Sergeyev2017}.
 
@@
 В результате такой редукции многомерная задача глобальной оптимизации (1) сводится к одномерной задаче
-In our research, we use a dimensionality reduction scheme based on Peano curves $y(x)$ that uniquely and continuously map the segment $[0,1]$ to an $N$-dimensional hypercube $D$ [REFS].
+In our research, we use a dimensionality reduction scheme based on Peano curves $y(x)$ that uniquely and continuously map the segment $[0,1]$ to an $N$-dimensional hypercube $D$ \cite{Strongin2000}.
 As a result of this reduction, the multidimensional global optimization problem (\ref{problem}) reduces to a one-dimensional problem
@@
 By using Peano curves, the boundedness of the function differences is preserved.
-If the original function $\phi(x)$ satisfied the Lipschitz condition, then the reduced one-dimensional function $\phi(y(x))$ from (\ref{reduced_problem}) will satisfy the Holder condition
+If the original function $\phi(x)$ satisfied the Lipschitz condition, then the reduced one-dimensional function $\phi(y(x))$ from (\ref{reduced_problem}) will satisfy the H{\"o}lder condition
 \begin{equation}\label{holder} 
@@
 График демонстрирует сложную структуру получающейся одномерной функции.
-In the numerical implementation of optimization algorithms, an evolvent is used to approximate the true Peano curve with a given accuracy, which is deter-mined by the desired accuracy of the search for the global extremum.
+In the numerical implementation of optimization algorithms, an evolvent is used to approximate the true Peano curve with a given accuracy, which is determined by the desired accuracy of the search for the global extremum.
 As an example, Fig.~\ref{fig1} shows a one-dimensional function resulting from the reduction of a two-dimensional problem.
@@
 
-\section{Parallel asynchronous global search algorithm}
+\section{Parallel asynchronous global search algorithm}\label{sec:PA}
 
@@
 Но в случае решения задач глобальной оптимизации такой подход будет неэффективным, поскольку при разделении области поиска только небольшая часть процессов будет работать в подобласти с искомым глобальным минимумом, все же остальные будут работать в подобластях, в которых отсутствует глобальный минимум исходной задачи.
-Various approaches to organizing parallel computations for various numerical algorithms have been developed to date [REFS].
+Various approaches to organizing parallel computations for various numerical algorithms have been developed to date \cite{Voevodin2015}.
 However, when solving multidimensional multiextremal optimization problems the known methods either do not provide proper effect or are not applicable altogether.
-For example, one of the main ways of organizing parallel computations consists in dividing the prob-lem solving domain between processes and solving the problem in parallel in these subdomains.
+For example, one of the main ways of organizing parallel computations consists in dividing the problem solving domain between processes and solving the problem in parallel in these subdomains.
 However, this approach will be inefficient for solving global optimization problems, because when dividing the search domain, only a small part of the processes will run in the subdomain with the desired global minimum, while the rest will run in the subdomains in which there is no global minimum of the original problem.
@@
 A more productive approach to parallelization is based on changing the sequential algorithm scheme for performing several trials in parallel.
-Note that in the problems of the class under consideration (problems of hyperparameter tun-ing for machine learning methods) a typical situation is when the objective function has different computational complexity at different points of the search domain.
+Note that in the problems of the class under consideration (problems of hyperparameter tuning for machine learning methods) a typical situation is when the objective function has different computational complexity at different points of the search domain.
 In such problems, the distribution of trial points between parallel processes should be performed asynchronously, on demand.
@@
 We used in our study an asynchronous global search algorithm, which implements a parallelizing scheme of the ``master/worker'' type.
-When describing the algorithm we will assume that we have at our disposal the following computational processp+1 : one master process andp worker processes.
+When describing the algorithm we will assume that we have at our disposal $p+1$ process: one master process and $p$ worker processes.
 The master process organizes the collection, accumulation and processing of search information (results of performed trials).
@@
 Правила работы процесса-мастера на любой из следующих итераций состоят в следующем.
-At the first iteration of the algorithm, the master process selects (uniformly or randomly) p different internal points of the search domain ${x^1, x^2, ..., x^p}, x^i \in (0,1), i=1, ..., p$, and distributes the trial points to the worker processes to perform trials in those points.
+At the first iteration of the algorithm, the master process selects (uniformly or randomly) $p$ different internal points of the search domain ${x^1, x^2, ..., x^p}, x^i \in (0,1), i=1, ..., p$, and distributes the trial points to the worker processes to perform trials in those points.
 The rules for the master process at any of the subsequent iterations are as follows.
@@
 Получив от рабочего процесса результаты испытания и обработав их мастер, в свою очередь, выбирает для него точку нового испытания x^(k+p+1) в соответствии со следующими правилами.
-Suppose $k$ tests have been performed, the master has received and processed their results, and during the current iteration the worker processes perform trials at the points of the set $S_k={x^(k+1), x^(k+2), …, x^(k+p)}$, i.e.
+Suppose $k$ tests have been performed, the master has received and processed their results, and during the current iteration the worker processes perform trials at the points of the set $S_k={x^{k+1}, x^{k+2},..., x^{k+p}}$, i.e.
 the trials have already started at these points but have not yet been completed.
 As soon as one of the worker processes completes a trial at some point (e.g.
-$x^(k+1)$), it forwards the results of the trial to the master process.
-After receiving the trial results from the worker process and processing them, the master process in turn selects a new trial point $x^(k+p+1)$ for the worker process according to the following rules.
+$x^{k+1}$), it forwards the results of the trial to the master process.
+After receiving the trial results from the worker process and processing them, the master process in turn selects a new trial point $x^{k+p+1}$ for the worker process according to the following rules.
 
@@
 %(рус.) 1. Перенумеровать в порядке возрастания все точки X_k  = {x^1,x^2,…,x^(k+p)  }, в которых либо проведены, либо проводятся испытания, т.е.
-\item Renumber, in ascending order, all the points $X_k={x^1,x^2,…,x^(k+p) }$ at which trials have either been carried out or are being carried out, i.e.
+\item Renumber, in ascending order, all the points $X_k={x^1,x^2,...,x^{k+p} }$ at which trials have either been carried out or are being carried out, i.e.
 \[
@@
 %(рус.) Теоретические основы алгоритма, а также примеры его применения приведены в работах [ссылки].
-The theoretical background of the algorithm as well as examples of its application are given in  [REFS].
+The theoretical background of the algorithm as well as examples of its application are given in  \cite{Barkalov2018,Barkalov2023,Gubaydullin2021,Strongin2020}.
 
-\section{Handling mixed parameters and functions that are not universally defined}
+\section{Handling mixed parameters and functions that are not universally defined}\label{sec:MP}
 
@@
 Let us number all different combinations of categorical parameters with integers $s$, i.e., let us assign to each number $s$ a vector $u_s$.
-Then (\ref{mixed_problem} ) can be rewritten as
+Then (\ref{mixed_problem}) can be rewritten as
 
@@
 %(рус.) Применяя к решению задачи (12) алгоритм глобального поиска, мы найдем решение задачи (10). 
-Applying the global search algorithm to solve problem (\ref{reduced_mixed_problem_2}), we will find a so-lution to problem (\ref{mixed_problem}).
+Applying the global search algorithm to solve problem (\ref{reduced_mixed_problem_2}), we will find a solution to problem (\ref{mixed_problem}).
 
@@
 А параметр 0<α<1 соответствует априорный оценке доли объема области поиска, в которой значения функции являются неопределенными.
-In the latter expression, $\alpha(1-1/r)^2\Delta_i$ corresponds to the lower estimate of the characteristic of the interval containing the current solution $z^*$ found from (\ref{cur_best_val}) (see [REF]) for the derivation of this estimate).
+In the latter expression, $\alpha(1-1/r)^2\Delta_i$ corresponds to the lower estimate of the characteristic of the interval containing the current solution $z^*$ found from (\ref{cur_best_val}) (see \cite{Strongin2020} for the derivation of this estimate).
 The parameter $0<\alpha<1$ corresponds to an a priori estimate of the portion of the search domain volume in which the function values are undefined.
@@
 
-\section{Results of computational experiments}
-
-Test links \cite{Seleznev2019,Blechschmidt2021,Xu2020,Strongin2000,Evtushenko2013,Evtushenko2009,Paulavicius2011,Gegrel2021,PaulaviciusZilinskas2014,Sergeyev2017,Usova2024,NIPS2015,Voevodin2015}
+\section{Results of computational experiments}\label{sec:RCE}
 
@@
 
-\noindent Displayed equations are centered and set on a separate line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
-
-%\begin{theorem}
-%This is a sample theorem. The run-in heading is set in bold, while
-%the following text appears in italics. Definitions, lemmas,
-%propositions, and corollaries are styled the same way.
-%\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-
 \begin{credits}
-\subsubsection{\ackname} This study was funded by the ...
+\subsubsection{\ackname} This study was supported by the Ministry of Science and Higher Education of the Russian Federation, project no. FSWR-2023-0034.
 

### 2025-04-09 2025/RSCD_MCO/RSD-2025.tex
Add grant number

@@
 %
-\authorrunning{S. Konnov, E. Kozinov, K. Barkalov, A. Sysoyev and V. Grishagin}
+\authorrunning{S. Konnov et al.}
 % First names are abbreviated in the running head.
@@
 
-Figure \ref{fig4} shows an example of built functions $R_{PS}(i)$ in the criteria domain.
+Fig.
+\ref{fig4} shows an example of built functions $R_{PS}(i)$ in the criteria domain.
 Red denotes the points assigned by the algorithm to the Pareto set.
@@
 \begin{credits}
-\subsubsection{\ackname} This study was funded by the ...
+\subsubsection{\ackname} The study  is supported by the program for the development of regional scientific and educational mathematical centers, Agreement No 075-02-2025-1727 with additional agreement No 075-02-2025-1727/1.
 

### 2025-04-10 2025/RuSCDays/RSD-2025.tex
correct one word
add results section and conclusion

@@
 \usepackage{hyperref}
+
+% добавили авторы
+\usepackage{multirow}
 
@@
 Поэтому применение для данных задач метаэвристических алгоритмов, требующих большого числа поисковых испытаний, является ограниченным.
-Более перспективным здесь является использование детерминированных алгорит-мов липшицевой оптимизации [Евтушенко, Сергеев, Стронгин], которые превосходят эвристические методы по качеству работы [Сергеев].
+Более перспективным здесь является использование детерминированных алгоритмов липшицевой оптимизации [Евтушенко, Сергеев, Стронгин], которые превосходят эвристические методы по качеству работы [Сергеев].
 Для методов липшицевой глобальной оптимизации предложен ряд подходов к распараллеливанию, обеспечивающих эффективную работу на суперкомпьютерных системах [Евтушенко, Жилинскас, Гергель].
@@
 %(рус.) Данная статья продолжает развитие подхода к параллельному решению задач липшицевой глобальной оптимизации, предложенного в работах [наши работы].
-Новым элементом является сочетание способов распараллеливания и предложенной в [Усова] схемы работы с частично вычислимыми функция-ми.
+Новым элементом является сочетание способов распараллеливания и предложенной в [Усова] схемы работы с частично вычислимыми функциями.
 Основная часть публикации имеет следующую структуру.
@@
 Section 3 содержит описание параллельного алгоритма глобального поиска.
-В Section 4 описан способ решения задач со смешанными (дискретными и непрерывными) па-раметрами, а также обсуждается способ работы с функциями, определенны-ми не во всей области поиска.
-Section 5 содержит результаты вычислитель-ных экспериментов, демонстрирующих эффективность предлагаемого под-хода при решении задачи оптимизации гиперпараметров методов прогнози-рования значений временного ряда.
+В Section 4 описан способ решения задач со смешанными (дискретными и непрерывными) параметрами, а также обсуждается способ работы с функциями, определенными не во всей области поиска.
+Section 5 содержит результаты вычислительных экспериментов, демонстрирующих эффективность предлагаемого подхода при решении задачи оптимизации гиперпараметров методов прогнозирования значений временного ряда.
 This paper continues to develop the approach to solving Lipschitz global optimization problems in parallel as proposed in [REFS ON OUR WORKS].
@@
 
-\section{Handling mixed parameters and functions that are not universally defined}\label{sec:MP}
+\section{Handling mixed parameters and functions that are not defined everywhere}\label{sec:MP}
 
@@
 
-%(рус.) Однако в рассматриваемых задачах настройки гиперпарметров целевая функция может быть не определена в некоторых (заранее неизвестных) точ-ках области поиска.
-Это может приводить возникновению интервалов, у ко-торых обе граничных точки имеют нулевой индекс, т.е.
+%(рус.) Однако в рассматриваемых задачах настройки гиперпарметров целевая функция может быть не определена в некоторых (заранее неизвестных) точках области поиска.
+Это может приводить возникновению интервалов, у которых обе граничных точки имеют нулевой индекс, т.е.
 значение целевой функции в них не определено.
-Возможный подход к обработке таких интер-валов состоит в использовании на шаге 4 алгоритма для вычисления характе-ристики вместо формулы (7) следующей формулы
+Возможный подход к обработке таких интервалов состоит в использовании на шаге 4 алгоритма для вычисления характеристики вместо формулы (7) следующей формулы
 However, in the hyperparameter tuning tasks under consideration, the objective function may be undefined in some (previously unknown) points in the search domain.
@@
 
-%(рус.) Формула (14) будет применяться, если значения как минимум в одной из граничных точек интервала значения целевой функции не определены, тогда как формула (15) будет использоваться только в случае, когда значения це-левой функции вычислены в обоих точках.
+%(рус.) Формула (14) будет применяться, если значения как минимум в одной из граничных точек интервала значения целевой функции не определены, тогда как формула (15) будет использоваться только в случае, когда значения целевой функции вычислены в обоих точках.
 Formula (\ref{new_point_1_2}) will be used if the values in at least one of the boundary points of the objective function value interval are undefined, while formula (\ref{new_point_2_2}) will be used only when the values of the objective function are calculated at both points.
@@
 
-\begin{table}
-\caption{Table captions should be placed above the tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
+%(рус.) Рассмотренные в статье методы реализованы в iOpt parallel solver, разрабатываемом в ННГУ им. Н.И. Лобачевского и предназначенном для решения многомерных задач глобальной оптимизации. 
+The methods considered in this paper are implemented in the iOpt parallel solver \cite{iOptURL} +REF!!!!!, which is being developed at Lobachevsky State University of Nizhni Novgorod  for solving multidimensional global optimization problems.
+
+%(рус.) В качестве модельной задачи нами была использована задача настройки гиперпараметров методов предсказания значений временного ряда (Time Series Forecasting).
+В качестве исходных данных был использован датасет montly beer prodaction [5.1], содержащий 476 valid data points.
+We used the task of hyperparameter tuning for Time Series Forecasting methods as a model problem.
+The monthly beer production dataset \cite{MonthlyBeerDataset,MonthlyBeerArticle} containing 476 valid data points was used as input data.
+
+%(рус.) Для решения задачи предсказания значений временного ряда мы использовали фреймворк машинного обучения FEDOT [5.2, 5.2.1] (open-source framework for automated modeling and machine learning (AutoML) problems).
+FEDOT поддерживает полный цикл решения задачи, включающий препроцессинг, выбор модели, настройку, кроссвалидацию и т.д.
+Фреймворк использует ML-модели, в основном из библиотек sklearn, statsmodels и keras, а также интегрирован с GOLEM open-source framework.
+To solve the problem of predicting time series values, we used the machine learning framework FEDOT \cite{FEDOT,nikitin2022automated} (open-source framework for automated modeling and machine learning problems).
+FEDOT supports the full problem solving cycle including preprocessing, model selection, tuning, cross validation, etc.
+The framework uses ML models, mainly from the sklearn, statsmodels and keras libraries, and is integrated with the GOLEM open-source framework.
+
+%(рус.) Построенный с использованием FEDOT pipeline включает в себя комбинацию из двух преобразований: lagged и cgru.
+Lagged-преобразование выполняется путем взятия значения переменной в предыдущий момент времени и включения его в модель в качестве признака в текущий момент времени.
+Для этого данные временного ряда сдвигаются на определенное количество шагов, которые называются лагом или временной задержкой.
+В качестве используемой нейронной сети будет выступать рекуррентная сеть (GRU), использующая свертку (CGRU).
+The pipeline constructed using FEDOT involves a combination of two transformations: lagged and cgru, Fig.~\ref{fig2}.
+Lagged transformation is performed by taking the value of a variable at a previous point in time and incorporating it into the model as a feature at the current point in time.
+For this purpose, the time series data is shifted by a certain number of steps, which are called  a lag  or time delay.
+A recurrent network (GRU) using convolution (CGRU) will serve as the neural network.
+
+\begin{figure}
+\center
+\includegraphics[width=0.5\textwidth]{fig2.png}
+\caption{Pipeline visualization.} \label{fig2}
+\end{figure}
+
+%(рус.) У lagged-преобразования имеется один параметр window_size, который изменяется от 5 до 250 с шагом 1.
+Поскольку этот параметр имеет большое количество значений, при оптимизации, будем рассматривать его как непрерывный (с округлением до единиц).
+The lagged transformation has one parameter $window\_size$, which varies from 5 to 250 in steps of 1.
+Since this parameter has a large number of values, we will consider it as continuous (rounded to units) during optimization.
+
+%(рус.) У cgru варьировались 4 дискретных параметра: cnn1_kernel_size  {4, 5}, cnn2_kernel_size  {5, 6}, cnn1_output_size  {32, 64}, cnn2_output_size  {32, 64}; и 2 непрерывных параметра: hidden_size  [20, 200], learning_rate  [0.0005, 0.005].
+У остальных параметров были зафиксированы default values: batch_size = 64, num_epochs = 50, optimizer = 'adamw', loss = ‘mse’.
+For cgru, 4 discrete and 2 continuous parameters were varied:
+\begin{itemize}
+\item cnn1\_kernel\_size $\in \{4, 5\}$,
+\item cnn2\_kernel\_size $\in \{5, 6\}$, 
+\item cnn1\_output\_size $\in \{32, 64\}$, 
+\item cnn2\_output\_size $\in \{32, 64\}$;
+\item hidden\_size $\in [20, 200]$, 
+\item learning\_rate $\in [0.0005, 0.005]$.
+\end{itemize}
+
+The other parameters had default values: batch\_size = 64, num\_epochs = 50, optimizer = 'adamw', loss = 'mse'.
+
+%(рус.) Таким образом в iOpt solver передавалась задача с тремя непрерывными параметрами и четырьмя дискретными параметрами, каждый из которых принимает 2 значения (всего 16 комбинаций).
+При некоторых сочетаниях параметров window_size, cnn1_kernel_size, cnn1_output_size, cnn2_kernel_size и cnn2_output_size невозможно вычислить значение критерия, т.е.
+метод возвращает бесконечное значение функции.
+Такие точки рассматривались как невычислимые.
+Thus a problem with three continuous parameters and four discrete parameters,  each of them taking  2 values (16 combinations in total), was submitted to the iOpt solver.
+With some combinations of parameters window\_size, cnn1\_kernel\_size, cnn1\_output\_size, cnn2\_kernel\_size and cnn2\_output\_size, it is impossible to compute the criterion value, as the method returns an infinite function value.
+Such points were regarded as as non-computable.
+
+%(рус.) Для демонстрации эффективности работы iOpt задача также была решена при помощи известного фреймворка оптимизации гиперпараметров Optuna [5.3, 5.3.1].
+Оба этих решателя были интегрированы в FEDOT и задействованы в вычислительном эксперименте.
+To demonstrate the effectiveness of iOpt, the problem was also solved using the well-known hyperparameter optimization framework Optuna \cite{OptunaURL,optuna}.
+Both of these solvers were integrated into FEDOT and employed in the computational experiment.
+
+%(рус.) Computations were conducted on the supercomputer “Lobachevsky” of the Nizhny Novgorod State University (operating system CentOS 7, control system SLURM, each supercomputer node had two processors Intel Xeon Silver 4310T CPU@2.30GHz (20 cores), 64Gb RAM).
+were used.
+Computations were performed using Python 3.9, FEDOT 0.7.5 [5.2, 5.2.1], iOpt  [20] and Optuna [5.3, 5.3.1].
+Calculations were performed on the ``Lobachevsky'' supercomputer of the Nizhni Novgorod State University (CentOS 7 operating system, SLURM control system, each supercomputer node had two Intel Xeon Silver 4310T processors CPU@2.30GHz (20 cores), 64Gb RAM).
+Data analyses were carried out using Python 3.9, FEDOT 0.7.5 \cite{FEDOT}, iOpt \cite{iOptURL} and Optuna \cite{OptunaURL}.
+
+%(рус.) Для решения поставленной задачи было выставлено ограничение на 1000 поисковых испытаний, число процессов P в параллельном алгоритме варьировалось от 1 до 20.
+В iOpt после 950 «глобальных» итераций глобального метода запускалось локальное уточнение с использованием Hooke-Jeeves method.
+To solve the problem, a limit of 1000 search trials was set, the number of processes P in the parallel algorithm varied from 1 to 20.
+In iOpt, after 950 ``global'' iterations of the global method, local refinement using the Hooke-Jeeves method was carried out.
+
+%(рус.) Результаты проведенного эксперимента отражены в таблице 1: указано время решения задачи и значение целевой метрики, полученное с помощью iOpt и OPTUNA.
+Число точек, в которых не получилось вычислить критерий оптимизации, составило порядка 10\% от общего числа точек search trials.
+Кривые, соответствующие полученному предсказанию, приведены на Fig.2.
+Темная линия соответствует истинным значениям временного ряда, голубая – предсказанным значениям, полученным после использования OPTUNA, желтая линия – предсказанным значениям, полученным после использования iOpt.
+The results of the experiment are shown in Table~\ref{tab1}: the time required to solve the problem and the value of the objective metric obtained using iOpt and Optuna.
+The number of points where the optimization criterion could not be calculated was about 10\% of the total number of search trials.
+The curves corresponding to the obtained prediction are shown in Fig.~\ref{fig3}.
+The dark line corresponds to the true values of the time series, the blue line corresponds to the predicted values obtained after using Optuna, and the yellow line corresponds to the predicted values obtained after using iOpt.
+
+\begin{figure}
+\center
+\includegraphics[width=1\textwidth]{fig3.png}
+\caption{Prediction visualization.} \label{fig3}
+\end{figure}
+
+\begin{table}[]
+\centering
+\caption{Results of solving the problem using iOpt and Optuna.}\label{tab1}
+\begin{tabular}{|l|c|c|c|c|}
 \hline
-Heading level &  Example & Font size and style\\
+\multirow{2}{*}{P} & \multicolumn{2}{c|}{iOpt} & \multicolumn{2}{c|}{Optuna} \\
+\cline{2-5} 
+ & Solving time & Final metric & Solving time & Final metric \\
 \hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
+1 & 6574,28 & 13,56 & 9554,08 & 13,43 \\
+2 & 3579,27 & 13,97 & 5127,50 & 13,69 \\
+4 & 1998,70 & 13,62 & 2598,68 & 13,26 \\
+8 & 1331,79 & 13,20 & 1275,45 & 13,26 \\
+16 & 846,69 & 13,31 & 1148,52 & 13,39 \\
+20 & 699,94 & 13,98 & 1578,67 & 13,32 \\
 \hline
@@
 \end{table}
+
+%(рус.) Ключевой характеристикой параллельных алгоритмов является ускорение.
+Мы будем использовать традиционное понятие ускорения как отношение времени последовательного запуска ко времени решения задачи в параллельном режиме.
+На Fig.3 приведены графики ускорения, полученного при параллельном запуске iOpt и OPTUNA.
+Наглядно видно, что iOpt на большом числе процессов показывает лучшее ускорение по сравнению с OPTUNA.
+The key characteristic of parallel algorithms is speedup.
+We will use the traditional notion of speedup as the ratio of the time of sequential running to the time of solving the problem in the parallel mode.
+Fig.~\ref{fig4} and Fig.~\ref{fig5} show the graphs of speedup obtained by parallel running of iOpt and Optuna.
+We can clearly see that iOpt shows better speedup on a large number of processes compared to OPTUNA.
+
+\begin{figure}
+\center
+\includegraphics[width=0.8\textwidth]{fig4.png}
+\caption{Speedup as a function of the number of processes.} \label{fig4}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width=0.8\textwidth]{fig5.png}
+\caption{Solving time depending on the number of processes.} \label{fig5}
+\end{figure}
+
+\section{Conclusion}
+
+%(рус.) Результаты экспериментов подтверждают, что алгоритм глобального поиска позволяет эффективно решать задачи настройки гиперпараметров методов искусственного интеллекта.
+Успешно обрабатываются ситуации, когда целевой критерий определен не всюду в области поиска.
+Параллельная программная реализация, выполненная в рамках iOpt solver, показывает хорошие ускорение и требует меньшего времени на решение задачи по сравнению с известными фреймворками для настройки гиперпараметров.
+Our experimental results confirm that the global search algorithm can effectively solve the problems of hyperparameter tuning for AI methods.
+Situations when the objective criterion is not defined everywhere in the search domain are also handled successfully.
+The parallel software implementation within the iOpt solver shows a good speedup and requires less time to solve the problem compared to known frameworks for hyperparameter tuning.
+
+%(рус.) Уменьшение прироста ускорения с ростом числа используемых процессов связана с необходимостью копирования большого объема данных, данная проблема будет решаться в дальнейшей работе. 
+%(рус.) Для распараллеливания вычислений использовалась библиотека multiprocess, которая хорошо себя показывает при передаче небольших объемов данных между процессами, однако при увеличении количества процессов объем данных, которые необходимо скопировать растет линейно, из-за чего наблюдаем замедление роста ускорения с увеличением числа используемых процессов.
+Дальнейшие исследования будут направлены на повышение ускорения и поиск более эффективного алгоритма копирования и переиспользования данных.
+To parallelize calculations in iOpt, we used the multiprocess library, which performs well when transferring small amounts of data between processes.
+However, when the number of parallel processes increases, the amount of data to be transferred between processes will grow linearly, which may cause the speedup to drop with the increasing number of processes being used.
+Further research will be geared towards  using a more efficient algorithm for copying and transferring data in iOpt in order to improve the speedup.
 

### 2025-04-11 2025/RuSCDays/RSD-2025.tex
pred-final version

@@
 Section 5 содержит результаты вычислительных экспериментов, демонстрирующих эффективность предлагаемого подхода при решении задачи оптимизации гиперпараметров методов прогнозирования значений временного ряда.
-This paper continues to develop the approach to solving Lipschitz global optimization problems in parallel as proposed in [REFS ON OUR WORKS].
+This paper continues to develop the approach to solving Lipschitz global optimization problems in parallel \cite{Barkalov2018,Gegrel2021,Strongin2018}.
 The new element is the combination of parallelization methods and the scheme for handling partially computable functions proposed in \cite{Usova2024}.
@@
 \begin{equation}\label{new_point_2} 
-x^{k+p+1} = \frac{x_t+x_{t-1}}{2}-\textit{sign}{(z_t-z_{t-1})}\frac{|z_t-z_{t-1}|^N}{2r\mu^N}, \text{ if } v(x_{t-1})=v(x_t).
+x^{k+p+1} = \frac{x_t+x_{t-1}}{2}-\frac{\textit{sign}{(z_t-z_{t-1})}}{2r} \left[ \frac{|z_t-z_{t-1}|}{\mu} \right]^N,
 \end{equation}
+if $v(x_{t-1})=v(x_t)$.
 %(рус.) Сразу после вычисления точки очередного испытания процесс-мастер добавляет ее в множество S и пересылает процессу-рабочему, который начинает проведение испытания в ней.
@@
 \begin{equation}\label{new_point_2_2} 
-x^{k+p+1} = \frac{x_t+x_{t-1}}{2}-\textit{sign}{(z_t-z_{t-1})}\frac{|z_t-z_{t-1}|^N}{2r\mu^N}, \text{ if } v(x_{t-1})=v(x_t) = 1.
+x^{k+p+1} = \frac{x_t+x_{t-1}}{2}-\frac{\textit{sign}{(z_t-z_{t-1})}}{2r} \left[ \frac{|z_t-z_{t-1}|}{\mu} \right]^N,
 \end{equation}
+if $v(x_{t-1})=v(x_t) = 1$.
 
@@
 %(рус.) Рассмотренные в статье методы реализованы в iOpt parallel solver, разрабатываемом в ННГУ им. Н.И. Лобачевского и предназначенном для решения многомерных задач глобальной оптимизации. 
-The methods considered in this paper are implemented in the iOpt parallel solver \cite{iOptURL} +REF!!!!!, which is being developed at Lobachevsky State University of Nizhni Novgorod  for solving multidimensional global optimization problems.
+The methods considered in this paper are implemented in the iOpt parallel solver \cite{iOptURL}, which is being developed at Lobachevsky State University of Nizhni Novgorod  for solving multidimensional global optimization problems.
 
@@
 В качестве исходных данных был использован датасет montly beer prodaction [5.1], содержащий 476 valid data points.
-We used the task of hyperparameter tuning for Time Series Forecasting methods as a model problem.
+We used the task of hyperparameter tuning for time series forecasting methods as a model problem.
 The monthly beer production dataset \cite{MonthlyBeerDataset,MonthlyBeerArticle} containing 476 valid data points was used as input data.
@@
 В качестве используемой нейронной сети будет выступать рекуррентная сеть (GRU), использующая свертку (CGRU).
-The pipeline constructed using FEDOT involves a combination of two transformations: lagged and cgru, Fig.~\ref{fig2}.
+The pipeline constructed using FEDOT involves a combination of two transformations: \textit{lagged} and \textit{cgru}, Fig.~\ref{fig2}.
 Lagged transformation is performed by taking the value of a variable at a previous point in time and incorporating it into the model as a feature at the current point in time.
@@
 У остальных параметров были зафиксированы default values: batch_size = 64, num_epochs = 50, optimizer = 'adamw', loss = ‘mse’.
-For cgru, 4 discrete and 2 continuous parameters were varied:
+For CGRU, 4 discrete and 2 continuous parameters were varied:
 \begin{itemize}
@@
 Calculations were performed on the ``Lobachevsky'' supercomputer of the Nizhni Novgorod State University (CentOS 7 operating system, SLURM control system, each supercomputer node had two Intel Xeon Silver 4310T processors CPU@2.30GHz (20 cores), 64Gb RAM).
-Data analyses were carried out using Python 3.9, FEDOT 0.7.5 \cite{FEDOT}, iOpt \cite{iOptURL} and Optuna \cite{OptunaURL}.
+Data analyses were carried out using Python 3.9, FEDOT 0.7.5 \cite{FEDOT}, iOpt 0.5.0 \cite{iOptURL} and Optuna 4.2.1 \cite{OptunaURL}.
 
@@
 
-\begin{table}[]
+\begin{table}
 \centering
@@
 \hline
-1 & 6574,28 & 13,56 & 9554,08 & 13,43 \\
-2 & 3579,27 & 13,97 & 5127,50 & 13,69 \\
-4 & 1998,70 & 13,62 & 2598,68 & 13,26 \\
-8 & 1331,79 & 13,20 & 1275,45 & 13,26 \\
-16 & 846,69 & 13,31 & 1148,52 & 13,39 \\
-20 & 699,94 & 13,98 & 1578,67 & 13,32 \\
+1 & 6574.2 & 13.5 & 9554.1 & 13.4 \\
+2 & 3579.2 & 13.9 & 5127.5 & 13.6 \\
+4 & 1998.7 & 13.6 & 2598.6 & 13.2 \\
+8 & 1331.7 & 13.2 & 1275.4 & 13.2 \\
+16 & 846.6 & 13.3 & 1148.5 & 13.3 \\
+20 & 699.9 & 13.9 & 1578.6 & 13.3 \\
 \hline
@@
 We will use the traditional notion of speedup as the ratio of the time of sequential running to the time of solving the problem in the parallel mode.
-Fig.~\ref{fig4} and Fig.~\ref{fig5} show the graphs of speedup obtained by parallel running of iOpt and Optuna.
-We can clearly see that iOpt shows better speedup on a large number of processes compared to OPTUNA.
+Fig.~\ref{fig4} and Fig.~\ref{fig5} show the graphs of speedup and time obtained by parallel running of iOpt and Optuna.
+We can clearly see that iOpt shows better speedup on a large number of processes compared to Optuna.
 

### 2025-04-14 2025/RuSCDays/RSD-2025.tex
Fig4 and Fig 5 are corrected

@@
 Получив от рабочего процесса результаты испытания и обработав их мастер, в свою очередь, выбирает для него точку нового испытания x^(k+p+1) в соответствии со следующими правилами.
-Suppose $k$ tests have been performed, the master has received and processed their results, and during the current iteration the worker processes perform trials at the points of the set $S_k={x^{k+1}, x^{k+2},..., x^{k+p}}$, i.e.
+Suppose $k$ trials have been performed, the master has received and processed their results, and during the current iteration the worker processes perform trials at the points of the set $S_k={x^{k+1}, x^{k+2},..., x^{k+p}}$, i.e.
 the trials have already started at these points but have not yet been completed.
@@
 In iOpt, after 950 ``global'' iterations of the global method, local refinement using the Hooke-Jeeves method was carried out.
+%\cite{Himmelblau72}
 

### 2025-04-14 2025/RuSCDays/RSD-2025.tex
some fix

@@
 
-\keywords{Global optimization  \and Multiextremal Functions \and Partially Defined Functions \and Hyperparameter Tuning \and  Time Series \and Artificial Intelligence.}
+\keywords{Global Optimization  \and Multiextremal Functions \and Partially Defined Functions \and Hyperparameter Tuning \and  Time Series \and Artificial Intelligence.}
 \end{abstract}
@@
 \begin{cases}
-    0     & \quad \text{if } x_i \text{ is boundary point},\\
-    1     & \quad \text{if } x_i \text{ is internal point}.
+    0,     & \quad \text{if } x_i \text{ is boundary point},\\
+    1,     & \quad \text{if } x_i \text{ is internal point}.
 \end{cases}
@@
 %(рус.) 4. Каждому интервалу (x_(i-1),x_i ),x_(i-1)  ∉ S_k,x_i  ∉ S_k,2≤ i≤ k+p, поставить в соответствие число R(i), которое называется характеристикой интервала и вычисляется по формуле
-\item Assign a number $R(i)$ to each interval $(x_{i-1}, x_i), x_(i-1) \notin S_k, x_i \notin S_k, 2 \leq i \leq k+p$, which is called the interval characteristic and is calculated using the formula
+\item Assign a number $R(i)$ to each interval $(x_{i-1}, x_i), x_{i-1} \notin S_k, x_i \notin S_k, 2 \leq i \leq k+p$, which is called the interval characteristic and is calculated using the formula
 \begin{equation}\label{characteristic}
@@
 Since these points are known in advance, they can be added to the information base of the algorithm at the first iteration with an index value equal to $0$.
-According to the rules of the algorithm, during its operation there will be no intervals with both boundary points having zero index, and the rules for calculating the features (\ref{characteristic}) and the points of the new interval (\ref{new_point_1})-(\ref{new_point_2}) will be valid.
+According to the rules of the algorithm, during its operation there will be no intervals with both boundary points having zero index, and the rules for calculating the characteristics (\ref{characteristic}) and the points of the new interval (\ref{new_point_1})-(\ref{new_point_2}) will be valid.
 
@@
 \begin{equation}\label{new_point_2_2} 
-x^{k+p+1} = \frac{x_t+x_{t-1}}{2}-\frac{\textit{sign}{(z_t-z_{t-1})}}{2r} \left[ \frac{|z_t-z_{t-1}|}{\mu} \right]^N,
+\begin{split}
+x^{k+p+1} = \frac{x_t+x_{t-1}}{2}-\frac{\textit{sign}{(z_t-z_{t-1})}}{2r} \left[ \frac{|z_t-z_{t-1}|}{\mu} \right]^N, \\
+\text{if }  v(x_{t-1})=v(x_t) = 1.
+\end{split}
 \end{equation}
-if $v(x_{t-1})=v(x_t) = 1$.
 
@@
 Поскольку этот параметр имеет большое количество значений, при оптимизации, будем рассматривать его как непрерывный (с округлением до единиц).
-The lagged transformation has one parameter $window\_size$, which varies from 5 to 250 in steps of 1.
+The lagged transformation has one parameter window\_size, which varies from 5 to 250 in steps of 1.
 Since this parameter has a large number of values, we will consider it as continuous (rounded to units) during optimization.

### 2025-06-09 2025/Problems of Informatics/article.tex
new introduction

@@
 \begin{center}
-	\textbf{АЛГОРИТМ РЕШЕНИЯ ЗАДАЧ ГЛОБАЛЬНОЙ ОПТИМИЗАЦИИ\break
-    С ЧАСТИЧНО ОПРЕДЕЛЕННЫМИ ФУНКЦИЯМИ}
+%Новое название
+	\textbf{АЛГОРИТМ ГЛОБАЛЬНОЙ ОПТИМИЗАЦИИ ДЛЯ НАСТРОЙКИ ГИПЕРПАРАМЕТРОВ МЕТОДОВ МАШИННОГО ОБУЧЕНИЯ}
 \end{center}
@@
 
-\textbf{Ключевые слова:} липшицева глобальная оптимизация, многоэкстремальные функции, функции вида <<черный ящик>>, частично определенные функции, скрытые ограничения, настройка гиперпараметров.
+\textbf{Ключевые слова:} машинное обучение, настройка гиперпараметров, глобальная оптимизация, функции вида <<черный ящик>>, частично определенные функции.
 \end{small}
@@
 \section{Введение}
-Глобальная оптимизация занимается разработкой методов, предназначенных для отыскания точек абсолютного (глобального) минимума многоэкстремальных функций.
-Возможность достоверной оценки глобального оптимума в таких задачах принципиально основана на наличии априорной информации о решаемой задаче, позволяющей связать возможные значения целевой функции с известными значениями в точках проведенных поисковых испытаний.
 
-Весьма часто такая информация представляется в виде предположения, что целевая функция $\phi(y)$ удовлетворяет условию Липшица с неизвестной априори константой $L$.
-Это предположение можно интерпретировать (применительно к прикладным задачам) как отражение ограниченности мощностей, порождающих изменения в моделируемой системе.
-При этом целевая функция зачастую задаётся в виде <<черного ящика>>, является недифференцируемой и каждое вычисление ее значения в некоторой точке допустимой области может требовать значительных вычислительных ресурсов.
+%новое введение
 
-Для решения задач липшицевой глобальной оптимизации разработан ряд эффективных детерминированных методов 
-[\ref{rfa:rulit:Jones2021},\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Birect2020},\ref{rfa:rulit:Sergeyev2017}].
-Проведенные сравнения показывают, что детерминированные алгоритмы превосходят (по разным критериям) широко распространенные природно-естественные алгоритмы [\ref{rfa:rulit:Liberti2005},\ref{rfa:rulit:Sergeyev2018}].
+В настоящее время методы искусственного интеллекта (ИИ) и машинного обучения (МО) применяются для решения широкого круга задач.
+К их числу можно отнести ставшие уже классическими задачи распознавания образов, классификации, регрессии.
+Однако в большинстве методов ИИ и МО имеются гиперпараметры, от выбора конкретных значений которых может зависеть качество (в некоторой метрике) полученного решения.
+Тот факт, что разница в качестве получаемого решения при разных значениях гиперпараметров может быть достаточно значительной, привел к возникновению класса задач настройки гиперпараметров (HPO --- hyperparameter optimization).
+С математической точки зрения такие задачи соответствуют задачам глобальной оптимизации с фиксированными границами изменения переменных.
+При этом часть гиперпараметров может быть категориальной, то есть принимать значения из некоторого дискретного множества, что приводит к необходимости использовать методы оптимизации, способные решать задачи с дискретными параметрами.
 
-Данная работа продолжает развитие одного из эффективных детерминированных методов решения задач липшицевой глобальной оптимизации -- ин\-фор\-ма\-ци\-он\-но-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}].
-Данный алгоритм предполагает решение многомерных задач посредством их сведения к задачам одномерной оптимизации при помощи кривых, заполняющих пространство (кривых Пеано).
+Основная сложность при решении  задач настройки гиперпараметров состоит  в  том, что поиск их оптимального сочетания требует для каждого выбранного варианта решения соответствующей исходной задачи машинного обучения, а значит, может быть довольно длительным по времени.
+Таким образом, любой подход будет ограничен в числе сочетаний гиперпараметров, которое метод глобальной оптимизации сможет проверить прежде, чем будет исчерпан доступный вычислительный ресурс.
 
-Отметим, что для оптимизации сложных объектов реального мира естественным является использование при расчетах сложных математических моделей, что, как следствие, существенно увеличивает трудоемкость поиска оптимума.
-За последние десятилетия специалистами в области оптимизации и параллельных вычислений было предложено множество способов снижения вычислительной сложности и ускорения используемых алгоритмов, связанных как с решением возникающих оптимизационных задач [\ref{rfa:rulit:Kvasov2013},\ref{rfa:rulit:Sergeyev2020}], так и с численным анализом исходных моделей [\ref{rfa:rulit:Dongarra2022},\ref{rfa:rulit:Duwe2020}].
+Весьма распространенными методами настройки параметров являются метаэвристические (генетические, имитационные и аналогичные им) алгоритмы [\ref{Zhou2021},\ref{Yang2022}].
+Данные алгоритмы широко применяются при отсутствии формульного описания оптимизируемой функции (функция вида <<черный ящик>>), что характерно для рассматриваемых задач.
+Метаэвристические алгоритмы слабо зависят от числа параметров задачи, используют информацию с предыдущих итераций для выполнения текущей, однако в силу заложенной в методы случайности дают гарантию отыскания глобального оптимума только в вероятностном смысле.
 
-Однако в последнее время актуальной становится принципиально новая проблема прикладных задач -- численная нестабильность исследуемых моделей в некоторых (заранее не известных) подобластях области изменения параметров.
-В таких подобластях невозможно корректно провести численное моделирование и вычислить значение целевой функции задачи.
-Данное явление можно интерпретировать или как наличие в задаче некоторых скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], или как наличие неизвестных областей, в которых целевая функция не является непрерывной [\ref{rfa:rulit:Audet2022}], или как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Sergeyev2003},\ref{rfa:rulit:Strongin2020}].
+Еще один подход к поиску глобального оптимума -- байесовская оптимизация, также применяемая для решения задач с функциями вида <<черный ящик>> [\ref{Frazier2018},\ref{Archetti2019}].
+Для своей работы методы байесовской оптимизации используют стохастическую модель оптимизируемой функции.
+Модель итерационно  обновляется на основе накапливаемой в процессе поиска оптимума информации, позволяя на каждой очередной итерации оценить наиболее вероятное положение глобального оптимума.
+Методы байесовской оптимизации обладают существенно  более  высокой  эффективностью  в  сравнении с полным перебором, но, тем не менее, также подвержены влиянию <<проклятия размерности>>, пусть и в меньшей степени.
+
+Хорошие результаты, достигаемые методами байесовской оптимизации, основаны на априорном предположении, что целевая функция задачи соответствует определенной стохастической модели, например, гауссовскому процессу.
+Однако в глобальной оптимизации существуют и иные предположения о виде функции, которые дают замечательные результаты.
+Одним  из таких  допущений о решаемой  задаче является предположение об ограниченности относительных изменений целевой функции.
+В этом случае говорят, что функция удовлетворяет условию Липшица, а решаемая задача называется задачей липшицевой глобальной оптимизации.
+Для класса задач липшицевой глобальной оптимизации разработан целый ряд эффективных алгоритмов [\ref{rfa:rulit:Jones2021},\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Birect2020},\ref{rfa:rulit:Sergeyev2017}], которые превосходят многие другие методы глобальной оптимизации [\ref{rfa:rulit:Liberti2005},\ref{rfa:rulit:Sergeyev2018}].
+
+В рассматриваемых задачах настройки гиперпараметров актуальной становится новая проблема --- некорректная работа исследуемых алгоритмов ИИ и МО в некоторых (заранее не известных) подобластях области изменения гиперпараметров.
+При этом программные реализации настраиваемых методов могут вести себя абсолютно по-разному: в лучшем случае информируют пользователя о том, что задача не может быть решена, в худшем возвращают недопустимое значение (NaN, inf) в качестве найденного решения.
+Проблема возникновения недопустимых комбинаций, которая не имеет особого значения в случае <<ручной>> настройки, стала камнем преткновения для многих фреймворков автоматической настройки гиперпараметров.
+Данное явление с математической точки зрения можно интерпретировать или как наличие в задаче некоторых скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], или как наличие неизвестных областей, в которых целевая функция не является непрерывной [\ref{rfa:rulit:Audet2022}], или как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Sergeyev2003},\ref{rfa:rulit:Strongin2020}].
 В такой постановке задача оптимизации существенно усложняется, т.к.
@@
 
-%НОВОЕ
-Характерным примером является задача настройки гиперпараметров методов искусственного интеллекта (ИИ) и машинного обучения (МО).
-Во многих случаях разница в эффективности (в заданной метрике) получаемого решения при разных значениях гиперпараметров может быть достаточно значительной.
-При этом вычислительная сложность настройки делает невозможным использование полного перебора.
-Данные особенности определили необходимость разработки различных методов интеллектуального автоматического подбора гиперпараметров.
-При этом проблема возникновения недопустимых комбинаций гиперпараметров, которая не имела особого значения в случае «ручной» настройки, для многих фреймворков, использующих методы интеллектуальной настройки, стала камнем преткновения.
-При этом сами настраиваемые методы могут вести себя абсолютно по-разному: в лучшем случае информируют пользователя о том, что задача не может быть решена, в худшем возвращают недопустимое значение (NaN, inf) в качестве найденного решения.
+%Отметим, что в распространенных фреймворках с открытым исходным кодом, предназначенных для решения задач оптимизации, отсутствует возможность работы с частично вычислимыми критериями.
 
-Отметим, что в распространенных фреймворках с открытым исходным кодом, предназначенных для решения задач оптимизации, отсутствует возможность работы с частично вычислимыми критериями.
-
+Настоящая работа продолжает развитие одного из эффективных детерминированных методов решения задач липшицевой глобальной оптимизации -- ин\-фор\-ма\-ци\-он\-но-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}].
 В статье приведено описание алгоритма глобального поиска, адаптированного для работы с частично определенной целевой функцией.
@@
 
-\section{Постановка задачи}
-В общем виде задача глобальной оптимизации может быть сформулирована следующим образом:
+\section{Постановка задачи настройки гиперпараметров}
+В общем виде задача настройки гиперпараметров (которой соответствует задача глобальной оптимизации) может быть сформулирована следующим образом:
 \begin{equation}\label{eq1} 
@@
 
+\item \label{Archetti2019}
+Archetti F, Candelieri A (2019) Bayesian Optimization and Data Science. SpringerBriefs in Optimization, Springer, Cham%   \doi{10.1007/978-3-030-24494-1}
+
+\item \label{Frazier2018}
+Frazier PI (2018) A tutorial on bayesian optimization. arXiv %  \doi{10.48550/arXiv.1807.02811}
+
 \end{enumerate}

### 2025-06-13 2025/RSCD_MCO/RSD-2025.tex
1. core -> kernel 2. fig2.png updeted

@@
 \includegraphics[width=0.6\textwidth]{fig2.png}
-\caption{Example of Pareto set evaluation in solving two scalar optimization problems.} \label{fig2}
+\caption{Example of Pareto set evaluation in solving two scalar optimization problems (values in the domain of criteria values).} \label{fig2}
 \end{figure}
@@
 \begin{enumerate}
-	\item When using the Support Vector Classification (SVC) model with a linear core, $R_{PS}(i)$ is considered on the base of the distance to the dividing hyperplane.
-	\item	When using the SVC model with an arbitrary core, $R_{PS}(i)$ RP is built according to probabilities of belonging trial points to the first class (Pareto set).
+	\item When using the Support Vector Classification (SVC) model with a linear kernel, $R_{PS}(i)$ is considered on the base of the distance to the dividing hyperplane.
+	\item	When using the SVC model with an arbitrary kernel, $R_{PS}(i)$ RP is built according to probabilities of belonging trial points to the first class (Pareto set).
 	\item Neural network models are used as a new way to evaluate $R_{PS}(i)$. The method of calculating $R_{PS}(i)$ is similar to step 2.
@@
 
-In the framework of the first approach, an SVC model with a linear core is built for computation $R_{PS}(i)$.
+In the framework of the first approach, an SVC model with a linear kernel is built for computation $R_{PS}(i)$.
 Information is extracted from the model to construct a dividing hyperplane equation between the selected point classes in the criteria value space.
@@
 	\item MGSA Dist -- the modified algorithm with computation $R_{PS}(i)$ from (\ref{eq:15}) based on distances,
-	\item MGSA Linear SVC, MGSA Poly SVC, MGSA Rbf SVC -- the algorithm modification with computation $R_{PS}(i)$ from (\ref{eq:15}) based on probabilities with different SVC algorithm core,
+	\item MGSA Linear SVC, MGSA Poly SVC, MGSA Rbf SVC -- the algorithm modification with computation $R_{PS}(i)$ from (\ref{eq:15}) based on probabilities with different SVC algorithm kernel,
 	\item MGSA NN -- the algorithm modification with computation $R_{PS}(i)$ from (\ref{eq:15}) on the base of probabilities obtained using neural networks.
@@
 Modifications of the algorithm can be distinguished that demonstrate the best results both in terms of the average number of iterations and in terms of the average HV.
-These include approaches using SVC (Linear, Rbf cores), a distance calculation approach, and neural networks with two hidden layers.
+These include approaches using SVC (Linear, Rbf kernels), a distance calculation approach, and neural networks with two hidden layers.
 Modification with one hidden layer shows slightly worse results.
-Modification  MGSA Poly SVC with polynomial core showed least improvement over MGSA.
+Modification  MGSA Poly SVC with polynomial kernel showed least improvement over MGSA.
 It is also worth noting that the modification of the MGSA Rbf SVC algorithm in the case of low accuracy demonstrates the best quality.

### 2025-06-13 2025/RSCD_MCO/RSD-2025.tex
1. поправил подпись fig.2 2. Добавил описание экспериментов раздел 5

@@
 \includegraphics[width=0.6\textwidth]{fig2.png}
-\caption{Example of Pareto set evaluation in solving two scalar optimization problems (values in the domain of criteria values).} \label{fig2}
+\caption{Example of Pareto set evaluation in solving two scalar optimization problems (trial values are presented in the criterion value range).} \label{fig2}
 \end{figure}
@@
 \ref{fig5} shows the dependence of the average value of the HV index on the average number of trials for the sequential algorithm.
+The number of $\lambda$ values from (\ref{eq:04}) in all experiments was equal to $\tau=50$.
+The $\lambda$ values were chosen to be uniformly distributed over the interval [0,1].
 Several solution accuracy values $\varepsilon = \{0.1,0.05,0.01\}$ were fixed for plotting the graph for algorithm (\ref{eq:13}).

### 2025-06-17 2025/RuSCDays/RSD-2025.tex
corrected according to reviewers comments

@@
 
+%
+%
+% !!!!!!!!!!!!!!! TO DO:: расширить обзор смежных исследований !!!!!!!!!!!!!
+% "мало внимания обзору существующих методов и подходов для решения задачи оптимизации гиперпараметров"
+%
+%
+
 %(рус.) Еще одной сложностью рассматриваемого класса задач HPO является наличие таких сочетаний значений гиперпараметров, при которых в некоторых (заранее не известных) подобластях области поиска исследуемый метод в принципе не работает.
@@
 By using Peano curves, the boundedness of the function differences is preserved.
-If the original function $\phi(x)$ satisfied the Lipschitz condition, then the reduced one-dimensional function $\phi(y(x))$ from (\ref{reduced_problem}) will satisfy the H{\"o}lder condition
+If the original function $\phi(y)$ satisfied the Lipschitz condition, then the reduced one-dimensional function $\phi(y(x))$ from (\ref{reduced_problem}) will satisfy the H{\"o}lder condition
 \begin{equation}\label{holder} 
@@
 
+%
+%
+% ЗАМЕЧАНИЕ: "не понятно назначение рисунка, т.к. читателю должен быть знаком общий вид одномерной функции"
+% Graph of the one-dimensional function after dimensionality reduction     of multiextremal objective function ????
+%
+%
+
 \begin{figure}
@@
 \includegraphics[width=0.8\textwidth]{fig1.png}
-\caption{Graph of the one-dimensional function after dimensionality reduction.} \label{fig1}
+\caption{Graph of the one-dimensional function after dimensionality reduction of multiextremal objective function.} \label{fig1}
 \end{figure}
@@
 \[
-0 = x_1 \leq x_2 \leq ... \leq x_{x+p} = 1,
+0 = x_1 \leq x_2 \leq ... \leq x_{k+p} = 1,
 \]
@@
 Если значение M получилось равным 0, то положить M=1.
-where $z_i=f(x_i)$, if $x_i \notin S_k, 1 \leq i \leq k+p$.
+where $z_i=\phi(y(x_i))$, if $x_i \notin S_k, 1 \leq i \leq k+p$.
 The values $z_i$ at the points $x_i \in S_k$ are undefined because the trials at these points have not yet been completed.
@@
 \begin{equation}\label{cur_best_val} 
-z^* = \min \left\{ \phi(x_i): 1 \leq i \leq k \right\}.
+z^* = \min \left\{ \phi(y(x_i)): 1 \leq i \leq k \right\}.
 \end{equation}
@@
 \begin{cases}
-    \Delta_i+\frac{(z_i-z_{i-1})^2}{(r\mu)^2\Delta_i} - 2 \frac{(z_i+z_{i-1}-2z^*)}{r\mu},   & \quad v(x_i)=v(x_{i-1})=1,\\
-    2\Delta_i- 4 \frac{(z_i-z^*)}{r\mu},   & \quad v(x_i)=1, v(x_{i-1})=0,\\
-    2\Delta_i- 4 \frac{(z_{i-1}-z^*)}{r\mu},       & \quad v(x_i)=0, v(x_{i-1})=1,\\
+    \Delta_i+\frac{(z_i-z_{i-1})^2}{(rM)^2\Delta_i} - 2 \frac{(z_i+z_{i-1}-2z^*)}{rM},   & \quad v(x_i)=v(x_{i-1})=1,\\
+    2\Delta_i- 4 \frac{(z_i-z^*)}{rM},   & \quad v(x_i)=1, v(x_{i-1})=0,\\
+    2\Delta_i- 4 \frac{(z_{i-1}-z^*)}{rM},       & \quad v(x_i)=0, v(x_{i-1})=1,\\
 \end{cases}
@@
 \begin{equation}\label{new_point_2} 
-x^{k+p+1} = \frac{x_t+x_{t-1}}{2}-\frac{\textit{sign}{(z_t-z_{t-1})}}{2r} \left[ \frac{|z_t-z_{t-1}|}{\mu} \right]^N,
+x^{k+p+1} = \frac{x_t+x_{t-1}}{2}-\frac{\textit{sign}{(z_t-z_{t-1})}}{2r} \left[ \frac{|z_t-z_{t-1}|}{M} \right]^N,
 \end{equation}
@@
 %(рус.) Задачу со смешанными параметрами можно свести к решению взаимосвязанного набора задач с непрерывными параметрами.
-Занумеруем целыми числами s все различные комбинации категориальных параметров, т.е.
+Занумеруем целыми числами все различные комбинации категориальных параметров, т.е.
 сопоставим каждому номеру s вектор u_s.
@@
 A problem with mixed parameters can be reduced to solving an interrelated set of problems with continuous parameters.
-Let us number all different combinations of categorical parameters with integers $s$, i.e., let us assign to each number $s$ a vector $u_s$.
+Let us number all different combinations of categorical parameters with integers, i.e., let us assign to each number $s$ a vector $u_s$.
 Then (\ref{mixed_problem}) can be rewritten as
@@
 \begin{equation}\label{mixed_problem_2} 
- \min \left\{\phi(y, u)\right\} = \min_{s \in \{1, ..., s\}} \left\{\phi(y, u_s): y \in D\right\}.
+ \min \left\{\phi(y, u)\right\} = \min_{s \in \{1, ..., S\}} \left\{\phi(y, u_s): y \in D\right\}.
 \end{equation}
@@
 \begin{equation}\label{reduced_mixed_problem} 
- \min \left\{\phi(y(x), u_s): x \in [0,1]\right\}, s \in \{1, ..., s\}.
+ \min \left\{\phi(y(x), u_s): x \in [0,1]\right\}, s \in \{1, ..., S\}.
 \end{equation}
@@
 \[
-f(x)=\phi(Y(x), uE(x) + 1), x \in [0, S],
+f(x)=\phi(Y(x), u_{E(x) + 1}), x \in [0, S],
 \]
@@
 \begin{cases}
-    \Delta_i+\frac{(z_i-z_{i-1})^2}{(r\mu)^2\Delta_i} - 2 \frac{(z_i+z_{i-1}-2z^*)}{r\mu},   & \quad v(x_i)=v(x_{i-1})=1,\\
-    2\Delta_i- 4 \frac{(z_i-z^*)}{r\mu},   & \quad v(x_i)=1, v(x_{i-1})=0,\\
-    2\Delta_i- 4 \frac{(z_{i-1}-z^*)}{r\mu},       & \quad v(x_i)=0, v(x_{i-1})=1,\\
+    \Delta_i+\frac{(z_i-z_{i-1})^2}{(rM)^2\Delta_i} - 2 \frac{(z_i+z_{i-1}-2z^*)}{rM},   & \quad v(x_i)=v(x_{i-1})=1,\\
+    2\Delta_i- 4 \frac{(z_i-z^*)}{rM},   & \quad v(x_i)=1, v(x_{i-1})=0,\\
+    2\Delta_i- 4 \frac{(z_{i-1}-z^*)}{rM},       & \quad v(x_i)=0, v(x_{i-1})=1,\\
     \alpha(1-1/r)^2\Delta_i,       & \quad v(x_i)=0, v(x_{i-1})=0,\\
@@
 \begin{split}
-x^{k+p+1} = \frac{x_t+x_{t-1}}{2}-\frac{\textit{sign}{(z_t-z_{t-1})}}{2r} \left[ \frac{|z_t-z_{t-1}|}{\mu} \right]^N, \\
+x^{k+p+1} = \frac{x_t+x_{t-1}}{2}-\frac{\textit{sign}{(z_t-z_{t-1})}}{2r} \left[ \frac{|z_t-z_{t-1}|}{M} \right]^N, \\
 \text{if }  v(x_{t-1})=v(x_t) = 1.
@@
 A recurrent network (GRU) using convolution (CGRU) will serve as the neural network.
+
+%
+%
+% TO DO: "возможно, вместе с pipeline, стоило бы проиллюстрировать архитектуру сети"
+%
+%
+%
 
@@
 Темная линия соответствует истинным значениям временного ряда, голубая – предсказанным значениям, полученным после использования OPTUNA, желтая линия – предсказанным значениям, полученным после использования iOpt.
-The results of the experiment are shown in Table~\ref{tab1}: the time required to solve the problem and the value of the objective metric obtained using iOpt and Optuna.
+%
+%
+%TO DO: "значение целевой метрики" - что за метрика использовалась в качестве целевой??? F_1????????
+%
+%
+The results of the experiment are shown in Table~\ref{tab1}: the time required to solve the problem and the value of the objective metric $F_1$ obtained using iOpt and Optuna.
 The number of points where the optimization criterion could not be calculated was about 10\% of the total number of search trials.
@@
 The dark line corresponds to the true values of the time series, the blue line corresponds to the predicted values obtained after using Optuna, and the yellow line corresponds to the predicted values obtained after using iOpt.
+
+%
+%
+%
+% TO DO:  "не ясно, какой вывод читатели должны сделать из этой иллюстрации. Не понятно, как оценивать качество прогнозирования из этого. Возможно, стоит на графике отобразить ошибку прогнозирования?"
+%
+%
+%
 
@@
 \end{figure}
-
+%
+%
+%
+%TO DO: указать единицу измерения времени (секунды?)
+%
+%
+%
 \begin{table}
@@
 \cline{2-5} 
- & Solving time & Final metric & Solving time & Final metric \\
+ & Solving time & Final $F_1$-score & Solving time & Final $F_1$-score \\
 \hline

### 2025-06-17 2025/RuSCDays/RSD-2025.tex
minor correction
Add sentence about Optuna and HyperOpt

@@
 there is a need to use global optimization methods capable of solving problems with mixed parameters.
+Such methods belonging to the class of metaheuristic and model-based algorithms are implemented in the well-known hyperparameter optimization frameworks Optuna \cite{optuna} and HyperOpt \cite{hyperopt}.
 
@@
 As soon as one of the worker processes completes a trial at some point (e.g.
-$x^{k+1}$), it forwards the results of the trial to the master process.
+$x^{k+1}$), it forwards the results of the trial to the master process and the point is excluded from the set $S_k$.
 After receiving the trial results from the worker process and processing them, the master process in turn selects a new trial point $x^{k+p+1}$ for the worker process according to the following rules.

### 2025-06-17 2025/RSCD_MCO/RSD-2025.tex
minor correction

--- 2025-06-13 2025/RSCD_MCO/RSD-2025.tex
@@
 \includegraphics[width=0.6\textwidth]{fig2.png}
-\caption{Example of Pareto set evaluation in solving two scalar optimization problems (values in the domain of criteria values).} \label{fig2}
+\caption{Example of Pareto set evaluation in solving two scalar optimization problems (trial points are presented in the criteria space).} \label{fig2}
 \end{figure}
@@
 
-As a criterion for the quality of the found Pareto set, the indicator HV \cite{Evtushenko2014,pymoo,AGP_ML} was applied.
-Higher value of the indicator corresponds to better solution.
+As a criterion for the quality of the found Pareto set, the indicator HV was applied. Higher value of the indicator corresponds to better solution. 
+HV index is a traditional quality metric when approximating a Pareto set \cite{Evtushenko2014,pymoo,AGP_ML}.
 To assess the performance of the algorithm, the number of trials carried out when solving the problem are given.
@@
 \ref{fig5} shows the dependence of the average value of the HV index on the average number of trials for the sequential algorithm.
+The number of $\lambda$ values from (\ref{eq:04}) was equal to $\tau=50$ in all experiments.
+The $\lambda$ values were chosen to be uniformly distributed over the interval [0,1].
 Several solution accuracy values $\varepsilon = \{0.1,0.05,0.01\}$ were fixed for plotting the graph for algorithm (\ref{eq:13}).
@@
 The position of the resulting curve for the chosen algorithm to the left and above corresponds to the best behavior of the algorithm as a whole.
-During the experiments the parameter $\alpha$ from (15) varied, Fig.
+During the experiments the parameter $\alpha$ from (15) varied within the range $[0.01,0.05]$ with a step of $0.01$, Fig.
 \ref{fig5} shows only the best results for each of the variants.

### 2025-06-17 2025/RSCD_MCO/RSD-2025.tex
fix rev. comments

@@
 %\urlstyle{rm}
-
+\usepackage{xcolor}
 \usepackage{amsmath}
@@
 
-\begin{figure}
+\begin{figure}[t]
 \center
-\includegraphics[width=0.7\textwidth]{fig1.png}
+\includegraphics[width=0.6\textwidth]{fig1.png}
 \caption{Computational scheme of the algorithm for solving multicriterial optimization problems using machine learning models.} \label{fig1}
@@
 There are different ways to build functions $F(\lambda_j,y)$ \cite{Miettinen1999,Konnov2025,Marler2004}.
+%В рамках предлагаемого алгоритма $\tau$ фиксируеться.
+Множество наборов $\lambda_j$, $1 \leq j \leq \tau$, выбирается случайно с использованием генератора Соболя в множестве  возможных значений (\ref{eq:051}).
+{\color{red} In the proposed algorithm, $\tau$ is fixed.
+The set of $\lambda_j$, $1 \leq j \leq \tau$, is chosen randomly using the Sobol generator\cite{Sobol2003} in the space of possible values (\ref{eq:051})}
+
+\begin{equation}
+\label{eq:051}
+\sum_{i=1}^s {\lambda_{i,j}} = 1, 0 \leq \lambda_{i,j} \leq 1.
+\end{equation}
+
 The performed computational experiments used minimax convolution of criteria
+
 \begin{equation}
 \label{eq:05}
-F(\lambda_j, y) = \max_{1 \leq i \leq s} \left(\lambda_{i,j} f_i (y)\right), y \in D, \sum_{i=1}^s {\lambda_{i,j}} = 1, 0 \leq \lambda_{i,j} \leq 1.
+F(\lambda_j, y) = \max_{1 \leq i \leq s} \left(\lambda_{i,j} f_i (y)\right), y \in D.
 \end{equation}
@@
 
-\begin{figure}
+\begin{figure}[t]
 \center
-\includegraphics[width=0.6\textwidth]{fig2.png}
+\includegraphics[width=0.7\textwidth]{fig2.png}
 \caption{Example of Pareto set evaluation in solving two scalar optimization problems (trial points are presented in the criteria space).} \label{fig2}
@@
 
-\begin{figure}
+\begin{figure}[t]
 \center
@@
 
-\begin{figure}
+\begin{figure}[t]
 \center
-\includegraphics[width=0.6\textwidth]{fig4.png}
+\includegraphics[width=\textwidth]{fig4.png}
 \caption{Examples of Pareto set evaluation for the case of two scalar optimization problems.} \label{fig4}
@@
 The algorithm proposed in the work was implemented in Python using the framework iOpt \cite{ioptmco}.
+%Сравнение предыдущей последовательной версии предлагаемого алгоритма с современными разработками можно найти в нашей статье [Konnov2025].
+В этом разделе больше внимания уделено анализу эффективности параллельной реализации.
+{\color{red}  A comparison of the previous sequential version of the proposed algorithm with state-of-the-art developments can be found in our paper \cite{Konnov2025}.
+This section focuses more on the analysis of the efficiently of parallel implementation.}
 
@@
 
-\begin{table}[ht]
+\begin{table}[t]
 \centering

### 2025-06-18 2025/RuSCDays/RSD-2025.tex
some fix

@@
 there is a need to use global optimization methods capable of solving problems with mixed parameters.
+%new
 Such methods belonging to the class of metaheuristic and model-based algorithms are implemented in the well-known hyperparameter optimization frameworks Optuna \cite{optuna} and HyperOpt \cite{hyperopt}.
@@
 For Lipschitz global optimization methods, a number of parallelization approaches have been proposed to ensure efficient operation in supercomputer systems \cite{Evtushenko2009,Gegrel2021,Paulavicius2011}.
-
-%
-%
-% !!!!!!!!!!!!!!! TO DO:: расширить обзор смежных исследований !!!!!!!!!!!!!
-% "мало внимания обзору существующих методов и подходов для решения задачи оптимизации гиперпараметров"
-%
-%
 
@@
 \includegraphics[width=0.8\textwidth]{fig1.png}
-\caption{Graph of the one-dimensional function after dimensionality reduction of multiextremal objective function.} \label{fig1}
+\caption{Graph of the one-dimensional function after dimensionality reduction of a multiextremal objective function.} \label{fig1}
 \end{figure}
@@
 Правила работы процесса-мастера на любой из следующих итераций состоят в следующем.
-At the first iteration of the algorithm, the master process selects (uniformly or randomly) $p$ different internal points of the search domain ${x^1, x^2, ..., x^p}, x^i \in (0,1), i=1, ..., p$, and distributes the trial points to the worker processes to perform trials in those points.
+At the first iteration of the algorithm, the master process selects (uniformly or randomly) $p$ different internal points of the search domain $\{x^1, x^2, ..., x^p\}, x^i \in (0,1), 1 \leq i \leq p$, and distributes the trial points to the worker processes to perform trials in those points.
 The rules for the master process at any of the subsequent iterations are as follows.
@@
 Получив от рабочего процесса результаты испытания и обработав их мастер, в свою очередь, выбирает для него точку нового испытания x^(k+p+1) в соответствии со следующими правилами.
-Suppose $k$ trials have been performed, the master has received and processed their results, and during the current iteration the worker processes perform trials at the points of the set $S_k={x^{k+1}, x^{k+2},..., x^{k+p}}$, i.e.
+Suppose $k$ trials have been performed, the master has received and processed their results, and during the current iteration the worker processes perform trials at the points of the set $W=\{x^{k+1}, x^{k+2},..., x^{k+p}\}$, i.e.
 the trials have already started at these points but have not yet been completed.
 As soon as one of the worker processes completes a trial at some point (e.g.
-$x^{k+1}$), it forwards the results of the trial to the master process and the point is excluded from the set $S_k$.
+$x^{k+1}$), it forwards the results of the trial to the master process and the point is excluded from the set $W$.
 After receiving the trial results from the worker process and processing them, the master process in turn selects a new trial point $x^{k+p+1}$ for the worker process according to the following rules.
@@
 %(рус.) 1. Перенумеровать в порядке возрастания все точки X_k  = {x^1,x^2,…,x^(k+p)  }, в которых либо проведены, либо проводятся испытания, т.е.
-\item Renumber, in ascending order, all the points $X_k={x^1,x^2,...,x^{k+p} }$ at which trials have either been carried out or are being carried out, i.e.
+\item Renumber, in ascending order, all the points $X_k=\{x^1,x^2,...,x^{k+p} \}$ at which trials have either been carried out or are being carried out, i.e.
 \[
@@
 \[
-M_1= \max \left\{ \frac{|z_i - z_{i - 1}|}{(x_i - x_{i - 1})^\frac{1}{N}}: x_{i-1} \notin S_k, x_i \notin S_k, 2 \leq i \leq k+p \right\},
+M_1= \max \left\{ \frac{|z_i - z_{i - 1}|}{(x_i - x_{i - 1})^\frac{1}{N}}: x_{i-1} \notin W, x_i \notin W, 2 \leq i \leq k+p \right\},
 \]
 \[
-M_2= \max \left\{ \frac{|z_{i+1} - z_{i - 1}|}{(x_{i+1} - x_{i - 1})^\frac{1}{N}}: x_{i} \in S_k, 2 \leq i < k+p \right\},
+M_2= \max \left\{ \frac{|z_{i+1} - z_{i - 1}|}{(x_{i+1} - x_{i - 1})^\frac{1}{N}}: x_{i} \in W, 2 \leq i < k+p \right\},
 \]
@@
 Если значение M получилось равным 0, то положить M=1.
-where $z_i=\phi(y(x_i))$, if $x_i \notin S_k, 1 \leq i \leq k+p$.
-The values $z_i$ at the points $x_i \in S_k$ are undefined because the trials at these points have not yet been completed.
+where $z_i=\phi(y(x_i))$, if $x_i \notin W, 1 \leq i \leq k+p$.
+The values $z_i$ at the points $x_i \in W$ are undefined because the trials at these points have not yet been completed.
 If the value $M$ turns out to be equal to $0$, then assume $M=1$.
@@
 %(рус.) 4. Каждому интервалу (x_(i-1),x_i ),x_(i-1)  ∉ S_k,x_i  ∉ S_k,2≤ i≤ k+p, поставить в соответствие число R(i), которое называется характеристикой интервала и вычисляется по формуле
-\item Assign a number $R(i)$ to each interval $(x_{i-1}, x_i), x_{i-1} \notin S_k, x_i \notin S_k, 2 \leq i \leq k+p$, which is called the interval characteristic and is calculated using the formula
+\item Assign a number $R(i)$ to each interval $(x_{i-1}, x_i), x_{i-1} \notin W, x_i \notin W, 2 \leq i \leq k+p$, which is called the interval characteristic and is calculated using the formula
 \begin{equation}\label{characteristic}
@@
 \[
-R(t)= \max \left\{ R(i): x_{i-1} \notin S_k, x_i \notin S_k, 2 \leq i \leq k+p\right\}.
+R(t)= \max \left\{ R(i): x_{i-1} \notin W, x_i \notin W, 2 \leq i \leq k+p\right\}.
 \]
@@
 %(рус.) Сразу после вычисления точки очередного испытания процесс-мастер добавляет ее в множество S и пересылает процессу-рабочему, который начинает проведение испытания в ней.
-Immediately after calculating the point of the next trial, the master process adds it to the set $S$ and forwards it to the worker process, which starts a trial there.
+Immediately after calculating the point of the next trial, the master process adds it to the set $W$ and forwards it to the worker process, which starts a trial there.
 \end{enumerate}
@@
 A problem with mixed parameters can be reduced to solving an interrelated set of problems with continuous parameters.
-Let us number all different combinations of categorical parameters with integers, i.e., let us assign to each number $s$ a vector $u_s$.
+Let us number all different combinations of categorical parameters with integers, i.e., let us assign to each number $s \in \{1, S\}$ a vector $u_s$.
 Then (\ref{mixed_problem}) can be rewritten as
@@
 %
-The results of the experiment are shown in Table~\ref{tab1}: the time required to solve the problem and the value of the objective metric $F_1$ obtained using iOpt and Optuna.
+The results of the experiment are shown in Table~\ref{tab1}: the time required to solve the problem and the value of the objective metric MSE obtained using iOpt and Optuna.
 The number of points where the optimization criterion could not be calculated was about 10\% of the total number of search trials.
@@
 \cline{2-5} 
- & Solving time & Final $F_1$-score & Solving time & Final $F_1$-score \\
+ & Solving time, sec. & Final MSE & Solving time, sec. & Final MSE \\
 \hline

### 2025-06-18 2025/RSCD_MCO/RSD-2025.tex
fix English in the comments

--- 2025-06-17 2025/RSCD_MCO/RSD-2025.tex
@@
 There are different ways to build functions $F(\lambda_j,y)$ \cite{Miettinen1999,Konnov2025,Marler2004}.
+In the proposed algorithm we used fixed value $\tau = 50$. The values $\lambda_j$, $1 \leq j \leq \tau$, are selected uniformly within the set of values satisfying the condition (\ref{eq:051})
+\begin{equation}
+\label{eq:051}
+\sum_{i=1}^s {\lambda_{i,j}} = 1, 0 \leq \lambda_{i,j} \leq 1.
+\end{equation}
+
 The performed computational experiments used minimax convolution of criteria
+
 \begin{equation}
 \label{eq:05}
-F(\lambda_j, y) = \max_{1 \leq i \leq s} \left(\lambda_{i,j} f_i (y)\right), y \in D, \sum_{i=1}^s {\lambda_{i,j}} = 1, 0 \leq \lambda_{i,j} \leq 1.
+F(\lambda_j, y) = \max_{1 \leq i \leq s} \left(\lambda_{i,j} f_i (y)\right), y \in D.
 \end{equation}
@@
 The algorithm proposed in the work was implemented in Python using the framework iOpt \cite{ioptmco}.
+% Результаты сравнения последовательной версии предлагаемого алгоритма с современными метаэвристическими решателями представлены в нашей статье \cite{Konnov2025}.
+Текущая работа сосредоточена на анализе эффективности параллельной версии алгоритма.
+The results of comparing the sequential version of the proposed algorithm with state-of-the-art metaheuristic solvers are presented in our paper \cite{Konnov2025}.
+The current work focuses on analyzing the efficiency of the parallel version of the algorithm.
 

### 2025-06-18 2025/RSCD_MCO/RSD-2025.tex
sort ref.

@@
 The criterion itself can often be presented not in an analytical form, but only as a procedure for calculating its values, i.e.
-in the form of a ``black box'' and additionally it can be multiextremal \cite{Miettinen1999,Ehrgott2005,Pardalos2017,Strongin2000,Sergeyev2013}.
+in the form of a ``black box'' and additionally it can be multiextremal \cite{Ehrgott2005,Miettinen1999,Pardalos2017,Sergeyev2013,Strongin2000}.
 For a more accurate description of the problem being solved, several (generally speaking, contradictory) optimized criteria can be taken.
@@
 When solving such problems using the scanning methods based on random or uniform grids, the number of calculations increases exponentially with the growth of dimensionality.
-It is possible to solve the considered problems using more efficient algorithms that can be divided into two large classes: deterministic \cite{Miettinen1999,Ehrgott2005,Pardalos2017,Konnov2025,Evtushenko2014} and non-deterministic (meta-heuristic) methods \cite{DPA02,Durillo2010,Mostaghim2007,Nebro2009,RC05,Zitzler2001}.
+It is possible to solve the considered problems using more efficient algorithms that can be divided into two large classes: deterministic \cite{Ehrgott2005,Evtushenko2014,Konnov2025,Miettinen1999,Pardalos2017} and non-deterministic (meta-heuristic) methods \cite{DPA02,Durillo2010,Mostaghim2007,Nebro2009,RC05,Zitzler2001}.
 
@@
 Deterministic algorithms guarantee an estimate of the globally optimal solution with a given accuracy, but in their basic implementations they can perform redundant calculations.
-To reduce the number of trials in this class of algorithms, heuristics are often used that reduce the amount of computations \cite{Konnov2025,AGP_ML}.
-Among effective deterministic algorithms for solving global optimization problems, the information-statistical family of global search methods \cite{Strongin2000,Sergeyev2013} can be considered.
-Initially, these algorithms were proposed for solving scalar optimization problems, subsequently, the algorithms were successfully developed to solve multicriterial problems of global search \cite{Konnov2025,MCO_ML_2023}.
+To reduce the number of trials in this class of algorithms, heuristics are often used that reduce the amount of computations \cite{AGP_ML,Konnov2025}.
+Among effective deterministic algorithms for solving global optimization problems, the information-statistical family of global search methods \cite{Sergeyev2013,Strongin2000} can be considered.
+Initially, these algorithms were proposed for solving scalar optimization problems, subsequently, the algorithms were successfully developed to solve multicriterial problems of global search \cite{MCO_ML_2023,Konnov2025}.
 
@@
 As a particular solution to the MCO problem, any effective option can be considered in which it is impossible to reduce the values of all criteria at once $f_i(y)$, $1 \leq i \leq s$, by selection of parameter values $y \in D$.
-In general case, when solving the MCO problem, it may be necessary to find the entire set of Pareto-optimal options \cite{Miettinen1999,Ehrgott2005,Pardalos2017}.
+In general case, when solving the MCO problem, it may be necessary to find the entire set of Pareto-optimal options \cite{Ehrgott2005,Miettinen1999,Pardalos2017}.
 
@@
 
-\textbf{Scalarization of efficiency criteria.} It is possible to find the Pareto set either solving the problem in the original formulation \cite{Evtushenko2014,DPA02,Durillo2010,Mostaghim2007,Nebro2009,RC05,Zitzler2001}, or reducing the original problem to a series of scalar problems to find effective solutions \cite{Pardalos2017,Konnov2025,Gergel2019_2,Gergel2018,GergelKozinov2020}.
+\textbf{Scalarization of efficiency criteria.} It is possible to find the Pareto set either solving the problem in the original formulation \cite{DPA02,Durillo2010,Evtushenko2014,Mostaghim2007,Nebro2009,RC05,Zitzler2001}, or reducing the original problem to a series of scalar problems to find effective solutions \cite{Gergel2019_2,Gergel2018,GergelKozinov2020,Konnov2025,Pardalos2017}.
 
@@
 Here $\lambda_j=\{\lambda_{i,j}\}$, $1 \leq i \leq s$, $1 \leq j \leq \tau$, are significance indicators for each criterion, s is the number of criteria, $\tau$ is the number of scalar optimization problems.
-There are different ways to build functions $F(\lambda_j,y)$ \cite{Miettinen1999,Konnov2025,Marler2004}.
+There are different ways to build functions $F(\lambda_j,y)$ \cite{Konnov2025,Marler2004,Miettinen1999}.
 In the proposed algorithm we used fixed value $\tau = 50$. The values $\lambda_j$, $1 \leq j \leq \tau$, are selected uniformly within the set of values satisfying the condition (\ref{eq:051})
@@
 
-\textbf{Dimensionality reduction.} For solving optimization problems (\ref{eq:04}) the dimensionality reduction on the base of Peano curves $y(x)$ that map unambiguously and continuously the segment $[0,1]$ to the $N$-dimensional hypercube $D$ is used \cite{Konnov2025,Gergel2019_2,GergelKozinov2020}.
+\textbf{Dimensionality reduction.} For solving optimization problems (\ref{eq:04}) the dimensionality reduction on the base of Peano curves $y(x)$ that map unambiguously and continuously the segment $[0,1]$ to the $N$-dimensional hypercube $D$ is used \cite{Gergel2019_2,GergelKozinov2020,Konnov2025}.
 As a result of this reduction, multidimensional global optimization problems (\ref{eq:04}) are reduced to one-dimensional subproblems
@@
 
-\textbf{Global search algorithm.} Information-statistical algorithm of global search (shortly GSA) \cite{Konnov2025,Gergel2019_2,Gergel2018,GergelKozinov2020,Strongin2000,Sergeyev2013} is used to solve problems (\ref{eq:06}).
+\textbf{Global search algorithm.} Information-statistical algorithm of global search (shortly GSA) \cite{Gergel2019_2,Gergel2018,GergelKozinov2020,Konnov2025,Sergeyev2013,Strongin2000} is used to solve problems (\ref{eq:06}).
 The computational scheme of GSA is as follows.
@@
 Thus, information about the calculated criteria is not lost when changing the task statement, but is reused.
-This approach has shown its effectiveness in solving many problems \cite{Konnov2025,Gergel2018,GergelKozinov2020}.
+This approach has shown its effectiveness in solving many problems \cite{Gergel2018,GergelKozinov2020,Konnov2025}.
 
@@
 An analysis of the impact of the selected model on the quality of the resulting solution is presented in Section \ref{sec4} with the results of computational experiments.
-Note that numerous computational experiments have shown that changing the method of calculating characteristic (\ref{eq:15}) leads to a significant increase in the efficiency of the algorithm as a whole \cite{Konnov2025,MCO_ML_2023}.
+Note that numerous computational experiments have shown that changing the method of calculating characteristic (\ref{eq:15}) leads to a significant increase in the efficiency of the algorithm as a whole \cite{MCO_ML_2023,Konnov2025}.
 
@@
 As a criterion for the quality of the found Pareto set, the indicator HV was applied. Higher value of the indicator corresponds to better solution. 
-HV index is a traditional quality metric when approximating a Pareto set \cite{Evtushenko2014,pymoo,AGP_ML}.
+HV index is a traditional quality metric when approximating a Pareto set \cite{AGP_ML,pymoo,Evtushenko2014}.
 To assess the performance of the algorithm, the number of trials carried out when solving the problem are given.

### 2025-06-22 2025/RuSCDays/RSD-2025.tex
delete todo comments
minor corrections

@@
 \end{itemize}
-
-%
-%
-% ЗАМЕЧАНИЕ: "не понятно назначение рисунка, т.к. читателю должен быть знаком общий вид одномерной функции"
-% Graph of the one-dimensional function after dimensionality reduction     of multiextremal objective function ????
-%
-%
 
@@
 
-%
-%
-% TO DO: "возможно, вместе с pipeline, стоило бы проиллюстрировать архитектуру сети"
-%
-%
-%
-
 \begin{figure}
@@
 В iOpt после 950 «глобальных» итераций глобального метода запускалось локальное уточнение с использованием Hooke-Jeeves method.
-To solve the problem, a limit of 1000 search trials was set, the number of processes P in the parallel algorithm varied from 1 to 20.
+To solve the problem, a limit of 1000 search trials was set, the number of processes $P$ in the parallel algorithm varied from 1 to 20.
 In iOpt, after 950 ``global'' iterations of the global method, local refinement using the Hooke-Jeeves method was carried out.
@@
 Темная линия соответствует истинным значениям временного ряда, голубая – предсказанным значениям, полученным после использования OPTUNA, желтая линия – предсказанным значениям, полученным после использования iOpt.
-%
-%
-%TO DO: "значение целевой метрики" - что за метрика использовалась в качестве целевой??? F_1????????
-%
-%
-The results of the experiment are shown in Table~\ref{tab1}: the time required to solve the problem and the value of the objective metric MSE obtained using iOpt and Optuna.
+
+The results of the experiment are shown in Table~\ref{tab1}: the time required to solve the problem and the MSE value obtained using iOpt and Optuna.
 The number of points where the optimization criterion could not be calculated was about 10\% of the total number of search trials.
@@
 The dark line corresponds to the true values of the time series, the blue line corresponds to the predicted values obtained after using Optuna, and the yellow line corresponds to the predicted values obtained after using iOpt.
-
-%
-%
-%
-% TO DO:  "не ясно, какой вывод читатели должны сделать из этой иллюстрации. Не понятно, как оценивать качество прогнозирования из этого. Возможно, стоит на графике отобразить ошибку прогнозирования?"
-%
-%
-%
 
@@
 \end{figure}
-%
-%
-%
-%TO DO: указать единицу измерения времени (секунды?)
-%
-%
-%
+
 \begin{table}

### 2025-07-07 2025/Problems of Informatics/article.tex
article after substantive revision

@@
 \usepackage{caption}
+
 \linespread{1.3}
@@
 % дополнительно
-%\usepackage{algorithm}
-%\usepackage{algpseudocode}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 \usepackage{multirow}
 \usepackage{listings}
+\usepackage{xcolor}
 
@@
 \begin{center}
-	\textbf{AN ALGORITHM FOR SOLVING GLOBAL OPTIMIZATION\break
-    PROBLEMS WITH PARTIALLY DINEFED FUNCTIONS}
+	\textbf{A GLOBAL OPTIMIZATION ALGORITHM FOR \break
+    TUNING HYPERPARAMETERS OF MACHINE LEARNING METHODS} 
 \end{center}
@@
 \begin{small}
+
 % аннотация англ.
+\colorbox{orange}{исправить в соответсвии с введением и экспериментами ->}
 The paper discusses the problem of finding a global minimum of a function that may be multiextremal, non-differentiable, given in the form of a ``black box'' and computable only in some part of the search domain.
@@
 \textbf{Keywords:} Lipschitz global optimization, multi-extremal functions, black-box functions, partially defined functions, hidden constraints, hyperparameter tuning.
+\colorbox{orange}{<-}
 \newline
@@
 %V.~A.~Grishagin, R.~A.~Israfilov, ``Global search acceleration in the nested optimization scheme'', AIP Conference Proceedings. {\bf 1738}, 400010 (2016). DOI: 10.1063/1.4952198.
+\item \label{rfa:enlit:Zhou2021}
+Zhou~J.
+, Qiu~Y.,  Zhu~S., Armaghani~D.
+J.
+, Li~C., Nguyen~H.
+, Yagiz~S.
+Optimization of support vector machine through the use of metaheuristic algorithms in forecasting {TBM} advance rate~// Eng.
+Appl.
+Artif.
+Intell.
+2021.
+Vol.~97.
+p.~104015.
+%DOI: 10.1016/j.engappai.2020.104015.
+
+\item \label{rfa:enlit:Yang2022}
+Yang~W., Xia~K., Fan~S., Wang~L., Li~T., Zhang~J., Feng~Y.
+A Multi-Strategy Whale Optimization Algorithm and Its Application~// Eng.
+Appl.
+Artif.
+Intell.
+2022.
+Vol.~108.
+p.~104558.
+%DOI: 10.1016/j.engappai.2021.104558.
+
+\item \label{rfa:enlit:Frazier2018}
+Frazier~P.I. A Tutorial on Bayesian Optimization~// arXiv. 2018. %DOI: 10.48550/arXiv.1807.02811.
+
+\item \label{rfa:enlit:Archetti2019}
+Archetti~F., Candelieri~A. Bayesian Optimization and Data Science. Cham: Springer Briefs in Optimization, 2019. %DOI: 10.1007/978-3-030-24494-1.
+
 \item \label{rfa:enlit:Jones2021}
-Jones~D., Martins~J. The direct algorithm: 25 years later~// J. Glob. Optim. 2021. Vol. 79, No,~3. P. 521--566. DOI: 10.1007/s10898-020-00952-6.
+Jones~D., Martins~J. The direct algorithm: 25 years later~// J. Glob. Optim. 2021. Vol.~79, No.~3. P.~521--566. %DOI: 10.1007/s10898-020-00952-6.
 
@@
 \item \label{rfa:enlit:PaulaviciusZilinskas2014}
-Paulavi{\v c}ius~R. and {\v Z}ilinskas~J. Simplicial Global Optimization. New York: Springer, 2014. DOI: 10.1007/978-1-4614-9093-7.
+Paulavi{\v c}ius~R. and {\v Z}ilinskas~J. Simplicial Global Optimization. New York: Springer, 2014. %DOI: 10.1007/978-1-4614-9093-7.
 
@@
 Paulavi{\v c}ius~R., Sergeyev~Y.D., Kvasov~D.E., {\v Z}ilinskas~J. Globally-biased {BIRECT} algorithm with local accelerators for expensive global optimization~// 
-Expert Syst. Appl. 2020. Vol. 144. p. 113052. DOI: 10.1016/j.eswa.2019.113052.
+Expert Syst. Appl. 2020. Vol.~144. p.~113052. %DOI: 10.1016/j.eswa.2019.113052.
 
@@
 \item \label{rfa:enlit:Liberti2005}
-Liberti~L., Kucherenko~S. Comparison of deterministic and stochastic approaches to global optimization~// Int. Trans. Oper. Res. 2005. Vol. 12. P. 263--285.
+Liberti~L., Kucherenko~S. Comparison of deterministic and stochastic approaches to global optimization~// Int. Trans. Oper. Res. 2005. Vol.~12. P.~263--285.
 
@@
 \item \label{rfa:enlit:Sergeyev2018}
-Sergeyev~Y.D., Kvasov~D.E., Mukhametzhanov~M.S. On the efficiency of nature-inspired metaheuristics in expensive global optimization with limited budget~// Sci. Rep. 2018. Vol. 8, No.~1. p. 435.
+Sergeyev~Y.D., Kvasov~D.E., Mukhametzhanov~M.S. On the efficiency of nature-inspired metaheuristics in expensive global optimization with limited budget~// Sci. Rep. 2018. Vol.~8, No.~1. p.~435.
 
@@
 \item \label{rfa:enlit:Sergeyev2013}
-Sergeyev~Y.D., Strongin~R.G., Lera~D. Introduction to Global Optimization Exploiting Space-Filling Curves. New York: Springer Briefs in Optimization, 2013. DOI: 10.1007/978-1-4614-8042-6.
+Sergeyev~Y.D., Strongin~R.G., Lera~D. Introduction to Global Optimization Exploiting Space-Filling Curves. New York: Springer Briefs in Optimization, 2013. %DOI: 10.1007/978-1-4614-8042-6.
 
@@
 \item \label{rfa:enlit:Kvasov2013}
-Kvasov~D.E., Sergeev~Ya.D. Metody lipshicevoj globalnoj optimizacii v zadachax upravleniya~// Avtomatika i telemexanika. 2013. No. 9. S. 3--19.
+Kvasov~D.E., Sergeev~Ya.D. Metody lipshicevoj globalnoj optimizacii v zadachax upravleniya~// Avtomatika i telemexanika. 2013. No.~9. S.~3--19.
 
@@
 Sergeyev~Y.D., Candelieri~A., Kvasov~D.E., Perego~R. Safe global optimization of expensive noisy black-box functions in the $\delta$-Lipschitz framework~// 
-Soft Comput. 2020. Vol. 24, No.~23. P. 17715--17735. DOI: 10.1007/s00500-020-05030-3.
+Soft Comput. 2020. Vol.~24, No.~23. P.~17715--17735. %DOI: 10.1007/s00500-020-05030-3.
 
@@
 \item \label{rfa:enlit:Dongarra2022}
-Dongarra~J.J. The evolution of mathematical software~// Commun. ACM. 2022. Vol. 65, No.~12. P. 66--72. DOI: 10.1145/3554977.
+Dongarra~J.J. The evolution of mathematical software~// Commun. ACM. 2022. Vol.~65, No.~12. P.~66--72. %DOI: 10.1145/3554977.
 
@@
 \item \label{rfa:enlit:Duwe2020}
-Duwe~K., et al. State of the art and future trends in data reduction for high-performance computing~// Supercomput. Front. Innov. 2020. Vol. 7, No.~1. DOI: 10.14529/jsfi200101.
+Duwe~K., et al. State of the art and future trends in data reduction for high-performance computing~// Supercomput. Front. Innov. 2020. Vol.~7, No.~1. %DOI: 10.14529/jsfi200101.
 
@@
 \item \label{rfa:enlit:Stripinis2021}
-Stripinis~L., Paulavi{\v c}ius~R. A new {DIRECT}-{GLh} algorithm for global optimization with hidden constraints~// Optim. Lett. 2021. Vol. 15, No.~6. P. 1865--1884.
-DOI: 10.1007/s11590-021-01726-z.
+Stripinis~L., Paulavi{\v c}ius~R. A new {DIRECT}-{GLh} algorithm for global optimization with hidden constraints~// Optim. Lett. 2021. Vol.~15, No.~6. P.~1865--1884.
+%DOI: 10.1007/s11590-021-01726-z.
 
@@
 \item \label{rfa:enlit:Audet2022}
-Audet~C., Batailly~A., Kojtych~S. Escaping unknown discontinuous regions in blackbox optimization~// SIAM J. Optim. 2022. Vol. 32, No.~3. P. 1843--1870. DOI: 10.1137/21m1420915.
+Audet~C., Batailly~A., Kojtych~S. Escaping unknown discontinuous regions in blackbox optimization~// SIAM J. Optim. 2022. Vol.~32, No.~3. P.~1843--1870. %DOI: 10.1137/21m1420915.
 
@@
 Candelieri~A. Sequential model based optimization of partially defined functions
-under unknown constraints~// J. Glob. Optim. 2019. Vol. 79, No.~2. P. 281--303. DOI: 10.1007/s10898-019-00860-4.
+under unknown constraints~// J. Glob. Optim. 2019. Vol.~79, No.~2. P.~281--303. %DOI: 10.1007/s10898-019-00860-4.
 
@@
 Strongin~R.G., Barkalov~K.A., Bevzuk~S.A. Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints~// 
-Soft Comput. 2020. Vol. 24, No.~16. P. 11853--11865. DOI: 10.1007/s00500-020-05078-1.
+Soft Comput. 2020. Vol.~24, No.~16. P.~11853--11865. %DOI: 10.1007/s00500-020-05078-1.
 
@@
 2024.
-Vol.
-1914.
-P.
-147--161.
-DOI: 10.1007/978-3-031-52470-7{\_}13.
+Vol.~1914.
+P.~147--161.
+%DOI: 10.1007/978-3-031-52470-7{\_}13.
 
@@
 Barkalov~K.A., et al. On solving the problem of finding kinetic parameters of catalytic isomerization of the pentane-hexane fraction using a parallel global search algorithm~// 
-Mathematics. 2022. Vol. 10, No.~19. p. 3665. DOI: 10.3390/math10193665.
+Mathematics. 2022. Vol.~10, No.~19. p.~3665. %DOI: 10.3390/math10193665.
 
@@
 2022.
-Vol.
-1618.
-P.
-293--306.
-DOI: 10.1007/978-3-031-11623-0{\_}20.
+Vol.~1618.
+P.~293--306.
+%DOI: 10.1007/978-3-031-11623-0{\_}20.
 
@@
 \item \label{rfa:enlit:differential_evolution}
-Storn~R., Price~K., Differential Evolution - a Simple and Efficient Heuristic for Global Optimization over Continuous Spaces // Journal of Global Optimization. 1997. Vol. 11. P. 341-359.
+Storn~R., Price~K., Differential Evolution - a Simple and Efficient Heuristic for Global Optimization over Continuous Spaces // Journal of Global Optimization. 1997. Vol.~11. P.~341-359.
 
 \item \label{rfa:enlit:dual_annealing}
-Xiang~Y, Gubian~S, Suomela~B, Hoeng~J. Generalized Simulated Annealing for Efficient Global Optimization: the GenSA Package for R // The R Journal. 2013. Vol. 5, No. 1.
+Xiang~Y, Gubian~S, Suomela~B, Hoeng~J. Generalized Simulated Annealing for Efficient Global Optimization: the GenSA Package for R // The R Journal. 2013. Vol.~5, No.~1.
 
 \item \label{rfa:enlit:direct}
-Gablonsky~J., Kelley~C. A Locally-Biased form of the DIRECT Algorithm // Journal of Global Optimization. 2001. Vol. 21. P. 27-37.
+Gablonsky~J., Kelley~C. A Locally-Biased form of the DIRECT Algorithm // Journal of Global Optimization. 2001. Vol.~21. P.~27--37.
 
@@
 1997.
-Vol.
-101.
-P.
-5111.
+Vol.~101.
+p.~5111.
 
@@
 Parallel Global Search Algorithm with Local Tuning for Solving Mixed-Integer Global Optimization Problems // Lobachevskii Journal of Mathematics.
-Vol.
-7.
-No.
-42.
+Vol.~7.
+No.~42.
 2021.
-P.
-1492-1503.
+P.~1492--1503.
 
@@
 2023.
-S.
-179-185.
+S.~179--185.
 
@@
 2003.
-Vol.
-29, No.~4.
-P.
-469--480.
+Vol.~29, No.~4.
+P.~469--480.
 
@@
 
-Весьма распространенными методами настройки параметров являются метаэвристические (генетические, имитационные и аналогичные им) алгоритмы [\ref{Zhou2021},\ref{Yang2022}].
+Весьма распространенными методами настройки параметров являются метаэвристические (генетические, имитационные и аналогичные им) алгоритмы [\ref{rfa:rulit:Zhou2021},\ref{rfa:rulit:Yang2022}].
 Данные алгоритмы широко применяются при отсутствии формульного описания оптимизируемой функции (функция вида <<черный ящик>>), что характерно для рассматриваемых задач.
@@
 
-Еще один подход к поиску глобального оптимума -- байесовская оптимизация, также применяемая для решения задач с функциями вида <<черный ящик>> [\ref{Frazier2018},\ref{Archetti2019}].
+Еще один подход к поиску глобального оптимума -- байесовская оптимизация, также применяемая для решения задач с функциями вида <<черный ящик>> [\ref{rfa:rulit:Frazier2018},\ref{rfa:rulit:Archetti2019}].
 Для своей работы методы байесовской оптимизации используют стохастическую модель оптимизируемой функции.
@@
 Последнее предположение делает невозможными применение известного ин\-фор\-ма\-ци\-он\-но-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска [\ref{rfa:rulit:Strongin2000}] или других методов липшицевой оптимизации [\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Sergeyev2017}].
-Для решения таких задач нами была разработана модификация алгоритма глобального поиска (АГП-Н), основанная на одном из предложенных ранее подходов [\ref{rfa:rulit:Usova2024}].
+Для решения таких задач нами была разработана модификация алгоритма глобального поиска, основанная на одном из предложенных нами ранее подходов [\ref{rfa:rulit:Usova2024}].
 
@@
 что позволяет применить для ее решения эффективные алгоритмы одномерной оптимизации.
-В качестве иллюстрации на Рис.~\ref{fig_peano} представлены образы отрезка $[0,1]$ для $N=2$, полученные с помощью разверток с плотностью $p=4,5$.
-
-\begin{figure}[h!]
-	\center{\includegraphics[scale=0.70]{figures/peano.pdf}}
-	\caption{Отображение отрезка $[0,1]$ с плотностью развертки: (а) $p=4$, (б) $p=5$}
-	\label{fig_peano}
-\end{figure}
+%В качестве иллюстрации на Рис.~\ref{fig_peano} представлены образы отрезка $[0,1]$ для $N=2$, полученные с помощью разверток с плотностью $p=4,5$.
+
+%\begin{figure}[h!]
+%	\center{\includegraphics[scale=0.70]{figures/peano.pdf}}
+%	\caption{Отображение отрезка $[0,1]$ с плотностью развертки: (а) $p=4$, (б) $p=5$}
+%	\label{fig_peano}
+%\end{figure}
 
@@
 
-\section{Алгоритм решения задач с частично определенной целевой функцией}
-
-Нашей целью является разработка алгоритма, в котором отсутствие значения целевой функции в некоторой точке не должно приводить к его некорректной работе. 
-Для этого нужно предложить новые правила обработки поисковых интервалов, в граничных точках которых отсутствуют значения целевой функции.
-Например, расчет характеристики интервала с двумя невычислимыми или с невычислимой и граничной точками можно производить, основываясь на длине интервала.
-Расчет характеристики интервала с невычислимой и вычислимой точками можно производить по правилам работы с граничным интервалом.
-
-Приведем подробное описание модификации алгоритма глобального поиска для не всюду вычислимой целевой функции (АГП-Н) для решения задач (\ref{eq2}).
-
-\subsection{Описание алгоритма АГП-Н} \label{algorithm_discr}
-
-На каждой итерации глобального поиска выполняется \textit{испытание}.
-Испытанием будем называть вычисление значения оптимизируемой функции $\phi (y(x))$ из (\ref{eq2}).
-Первое испытание будет осуществляться в серединной внутренней точке $x^1 \in (0,1)$.
-Выбор точки $x^{k+1}, k \geq 1$ очередного $(k+1)^\text{-го}$ испытания осуществляется на основе следующих правил.
-
-\textit{Правило 1.} Перенумеровать (нижним индексом) точки $x^i, 0 \leq i \leq k$ предшествующих испытаний в порядке возрастания значений координаты, то есть
-\begin{equation}\label{eq5} 
-0=x_0 < x_1 < ... < x_i < ... < x_{k}=1
-\end{equation}
-и сопоставить им значения $z_i=f(x_i), 0 < i < k$, вычисленные в этих точках, и индекс $v_i=v(x_i)$, определяемый по правилу
+\section{Алгоритм решения задач с частично определенной целевой функцией}\label{alg_discr}
+
+Алгоритм глобального поиска предполагает построение последовательности точек $y^i$, так называемых, \textit{поисковых испытаний}, для которых вычисляются значения целевой функции $z^i = \varphi(y^i)$.
+Согласно используемой схеме редукции размерности, проведение испытания также требует вычисление значения $y^i=y(x^i)$.
+При этом каждой точке испытания $x^i$ ставится в соответствие индекс $v^i$, определяемый по правилу
 \begin{equation}\label{eq6} 
-v_i=v(x_i)=
+v(x^i) =
   \begin{cases}
-    -1, & {\quad \text{если } x_i \text{ -- граничная точка}},\\
-    0, & {\quad \text{если } x_i \text{ -- невычислимая точка}},\\
-    1, & {\quad \text{если } x_i \text{ -- внутренняя точка}}.
+    -1, & {\quad \text{если } x^i \text{ -- граничная точка}},\\
+    0, & {\quad \text{если } x^i \text{ -- невычислимая точка}},\\
+    1, & {\quad \text{если } x^i \text{ -- внутренняя точка}}.
   \end{cases}
 \end{equation}
-Точки $x_0=0$ и $x_{k}=1$ введены дополнительно (значения $z_0$ и $z_{k}$ не определены).
-
-\textit{Правило 2.} Вычислить текущую нижнюю оценку
-\begin{equation}\label{eq7} 
-\mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i},\quad v(x_{i-1})=v(x_i)=1, \quad 2 \leq i \leq k  \right\},\quad \Delta _i= (x_i-x_{i-1})^{1/N},
-\end{equation}
-константы Гёльдера $K$ из (\ref{eq4}) для редуцированной функции $f(x)$, где $r > 1$ -- параметр надёжности. Если $\mu = 0$, то принять $\mu = 1$.
-
-\textit{Правило 3.} Определить текущее лучшее значение целевой функции
-\begin{equation}\label{eq8} 
-z^*= \min \left\{ f(x_i): v(x_i)=1, 0 < i < k \right\}.
-\end{equation}
-
-\textit{Правило 4.} Для каждого интервала $(x_{i-1},x_i),1 \leq i \leq k$, вычислить значение $R(i)$, называемое \textit{характеристикой} интервала, согласно выражению
-\begin{equation}\label{eq9} 
-R(i)=
-  \begin{cases}
-    \alpha{(1-\frac{1}{r})}^2 \Delta _i,& {\quad v(x_i)=v(x_{i-1})=0 \text{ или } v(x_i)=0 \text{ и } v(x_{i-1})=-1}\\
-    {} & {\quad \text{или } v(x_i)=-1  \text{ и } v(x_{i-1})=0},\\
-    \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1},\\
-    2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad  v(x_{i-1})=-1 \text{ или } v(x_{i-1})=0, v(x_i)=1},\\
-    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1 \text{ или } v(x_i)=0,}
-  \end{cases}
-\end{equation}
-где $\alpha < 1.0$ - задаваемый параметр метода.
-
-\textit{Правило 5.} Выбрать интервал $(x_{t-1},x_t)$ с максимальным значением характеристики $R(i)$: $R(t)= \max\{R(i): \; 1 \leq i \leq k\}$.
-Если максимальную характеристику имеют несколько интервалов, выбрать интервал %с большим количеством вычислимых точек
-с минимальным номером.
-
-\textit{Правило 6.} Провести очередное испытание в серединной точке интервала $(x_{t-1},x_t)$, если индексы его концевых точек не совпадают или оба соответствуют невычислимым точкам, т.е.
-\begin{equation}\label{eq10} 
-x^{k+1}=\frac {x_t+x_{t-1}}{2},\quad v(x_{t-1})\neq v(x_t) \text{ или } v(x_{t-1})=v(x_t)=0.
-\end{equation}
-В противном случае провести испытание в точке 
-\begin{equation}\label{eq11} 
-x^{k+1}= \frac {x_t+x_{t-1}}{2} -  \text{sign} {(z_t-z_{t-1})} \frac{1}{2r} \left[\frac {{|z_t-z_{t-1}|}}{\mu} \right]^N.
-\end{equation}
-
-\textit{Условие остановки.} Поиск завершен, если длина интервала $(x_{t-1},x_t)$, содержащего очередную точку $x^{k+1}$, не превышает заданной точности $\varepsilon$, то есть $\Delta _t \leq \varepsilon$, а одна из граничных точек данного интервала является внутренней точкой, т.е.
-$\min{\{v_{t-1}, v_t\}} = 1$, где $t$ из Правила 5, а $\varepsilon>0$ -- задаваемый параметр алгоритма.
+
+Результатом испытания является набор значений $(x^i, y^i=y(x^i), z^i = \varphi(y^i), v^i = v(x^i))$.
+
+Накопленная после проведения $k$ испытаний поисковая информация хранится в порядке возрастания значений координаты $x^i$ в множестве $\Omega_k$, в котором соответствующие записи для удобства перенумерованы нижним индексом.
+
+На каждом шаге алгоритма производится вычисление оценки константы Гёльдера, определение характеристик поисковых интервалов, выбор наиболее <<перспективного>> интервала и вычисление точки $y^{k+1}$ для проведения очередного испытания.
+
+Основные изменения в алгоритме затронули правило вычисления характеристики интервала (Алгоритм \ref{charact_alg}).
+Сходимость метода в этом случае требует доказательства, которое будет продемонстировано авторами в будущих статьях.
 
 %\begin{algorithm}
-%\caption{Алгоритм глобального поиска}\label{alg:cap}
-%\begin{algorithmic}
-%\Require $\varphi(y), D, r, \epsilon, K_{max}$
-%\Ensure $y_{min}$
-%\State $k = 2$
-%\State $t = 2$
-%\State $\Omega_k = \left\{ (x_1 = 0, y_1 = y(x_1), z_1 = \varphi(y_1)), (x_2 = 1, y_2 = y(x_2), z_2 = \varphi(y_2)) \right\}$
-%\While{ $\Delta_t \geq \epsilon \text{ and } k \leq K_{max}$}
-%    \State Compute the estimate of the H\"older constant $\mu$ according to (\ref{eq7})
-%    \For{$i=1$ \textbf{to} $k$}
-%        \State Compute the characteristic $R(i)$ according to (\ref{eq9})
-%    \EndFor
-%    \State Determine the number $t$, for which $R(t) = \max \left\{ R(i), 1 \leq i \leq k \right\}$
-%    \State Compute the new trial point $x^{k+1}$ according to (\ref{eq10}-\ref{eq11})
-%    \State Perform a new trial at the point $x^{k+1}$
-%    \State Add $(x^{k+1}, y^{k+1}, z^{k+1})$ to the set $\Omega_k$
-%    \State $k=k + 1$
-%\EndWhile
-%\State $l = \arg \min \left\{ z_i, 1 < i \leq k \right\}$
-%\State  $y_{min} = y_l$\\
-%\Return $y_{min}$
+%\caption{Алгоритм оценки константы Гёльдера }\label{mu_alg}
+%\begin{algorithmic}[1]
+%\Require $\Omega_k$
+%\Ensure $M$
+%\State $M \gets \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i},\quad v(x_{i-1}) = 1 \text{ and } v(x_i) = 1, \quad 1 \leq i \leq k  \right\}$
+%    \If {$M = 0$}
+%       \State $M = 1$
+%   \EndIf\\
+%\Return M
 %\end{algorithmic}
@@
 
-\textbf{Замечание 1 (о существовании вычислимого отрезка на кривой Пеано)}.
-В случае попадания точек испытаний в область невычислимости на первых итерациях, алгоритм будет выбирать лучший интервал на основе длин интервалов (т.е.
-будет строится равномерная сетка).
-Так как мы предполагаем, что область вычислимости целевой функции $Q = D \backslash I$ -- положительного объёма, то последующее сгущение сетки приведет к попаданию очередной точки испытания в область $Q$.
-
-%\begin{figure}[h!]
-%	\center{\includegraphics[scale=0.30]{figures/fig_0.png}}
-%	\caption{Пример расстановки точек испытаний до момента попадания в вычислимую точку}
-%	\label{fig_1}
-%\end{figure}
+\begin{algorithm}
+\caption{Алгоритм определения характеристики интервала}\label{charact_alg}
+\begin{algorithmic}[1]
+\Require $\Omega_k, i, z^*, \mu, r >1, 0 < \alpha \leq 1$
+\Ensure $R$
+
+\If {$v(x_{i-1}) = 1 \text{ and } v(x_{i}) = 1$}  \Comment{ обе точки являются внутренними }
+           \State $R \gets \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {(z_i+z_{i-1}-2z^*)}{r \mu}$
+\ElsIf {$v(x_{i}) = 1$}  \Comment{ точка $x_i$ является внутренней }
+           \State $R \gets 2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}$
+\ElsIf {$v(x_{i-1}) = 1$} \Comment{ точка $x_{i-1}$ является внутренней }
+            \State $ R \gets 2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}$
+\Else \Comment{ обе точки явлются невычислимыми или граничными }
+           \State $R \gets \alpha{(1-\frac{1}{r})}^2 \Delta _i$
+\EndIf\\
+\Return R
+\end{algorithmic}
+\end{algorithm}
+
+\textit{Примечание}. Здась и далее $\Delta_i$ используется для обозначаения длины $i$-го интервала в N-мерном пространстве, т.е.
+\[
+\Delta _i= (x_i-x_{i-1})^{1/N}.
+\]
+
+Подробное описание правил алгоритма глобального поиска на случай не всюду вычислимой целевой функции (АГП-Н)  для решения задач (\ref{eq2}) приведено в Алгоритме \ref{main_alg}.
+
+\begin{algorithm}
+\caption{Алгоритм глобального поиска на случай не всюду вычислимой целевой функции (АГП-Н)}\label{main_alg}
+\begin{algorithmic}[1]
+\Require $\varphi(y), D, \varepsilon, K_{max}, r >1, 0 < \alpha \leq 1$
+\Ensure $y^{min}$
+\State $\Omega_k \gets \left\{(x^1 = 0.0, y^1 = y(x^1), v^1=v(x^1)), (x^2 = 1.0, y^2 = y(x^2), v^2 = v(x^2))\right\}$
+\State $\Omega_k \gets \Omega_k \cup \left\{(x^3 = 0.5, y^3 = y(x^3), z^3 = \varphi(y^3), v^3 = v(x^3))\right\}$
+\State $z^* \gets z^3; k \gets 3; t \gets 2$
+\While{ $\Delta_t \geq \varepsilon \text{ and } k \leq K_{max} \text{ or } v(x_t) \neq 1 \text{ and } v(x_{t-1}) \neq 1 $}
+    \State $\mu \gets \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i},\quad v(x_{i-1}) = 1 \text{ and } v(x_i) = 1, \quad 1 \leq i \leq k  \right\}$
+    \If {$\mu = 0$}
+       \State $\mu = 1$
+   \EndIf
+    \For{$i=1$ \textbf{to} $k$}
+        \State $R_i \gets R(\Omega_k, i, z^*, \mu, r, \alpha)$
+    \EndFor
+    \State $t \gets \min \left\{\arg\max \left\{ R_i, 1 \leq i \leq k \right\}\right\}$
+     \If {$v(x_{t-1}) = 1 \text{ and } v(x_t) = 1$}
+        \State $x^{k+1} \gets \frac {x_t+x_{t-1}}{2} -  \text{sign} {(z_t-z_{t-1})} \frac{1}{2r} \left[\frac {{|z_t-z_{t-1}|}}{\mu} \right]^N$
+    \Else
+        \State $x^{k+1} \gets \frac {x_t+x_{t-1}}{2}$
+    \EndIf
+    \State $y^{k + 1} \gets y(x^{k + 1}); z^{k + 1} \gets \varphi(y^{k + 1}); v^{k+1} \gets v(x^{k+1})$
+    \State $z^* = \min \left \{z^*, z^{k+1}\right \}$
+    \State$\Omega_k \gets \Omega_k \cup \left\{(x^{k+1}, y^{k+1}, z^{k+1}, v^{k+1})\right\}$
+    \State $k \gets k + 1$
+\EndWhile
+\State $min \gets \arg \min \left\{ z^i, 1 < i < k \right\}$\\
+\Return $y^{min}$
+\end{algorithmic}
+\end{algorithm}
+
+\textbf{Замечание 1 (о связи с АГП)}.
+В случаях, когда задача не содержит скрытых ограничений или проводимые алгоритмом испытания не попадают в существующие области невычислимости, АГП-Н полностью повторяет работу своего прототипа -- алгоритма глобального поиска.
 
@@
 
-\textbf{Замечание 3 (о связи с АГП)}.
-В случаях, когда задача не содержит скрытых ограничений или проводимые алгоритмом испытания не попадают в существующие области невычислимости, АГП-Н полностью повторяет работу своего прототипа -- алгоритма глобального поиска.
+\textbf{Замечание 3 (о существовании вычислимого отрезка на кривой Пеано)}.
+В случае попадания точек испытаний в область невычислимости на первых итерациях, алгоритм будет выбирать лучший интервал на основе длин интервалов (т.е.
+будет строится равномерная сетка).
+Так как мы предполагаем, что область вычислимости целевой функции $Q = D \backslash I$ -- положительного объёма, то последующее сгущение сетки приведет к попаданию очередной точки испытания в область $Q$.
+
+%\begin{figure}[h!]
+%	\center{\includegraphics[scale=0.30]{figures/fig_0.png}}
+%	\caption{Пример расстановки точек испытаний до момента попадания в вычислимую точку}
+%	\label{fig_1}
+%\end{figure}
 
@@
 
-%\subsection{Схема распараллеливания}
-
-%Пусть выполнено $n \geq 1$ итераций метода, в процессе которых были проведены испытания в $k(n)$ точках $x^i,1 \leq i \leq k$.
-Тогда точки $x^{k+1},…,x^{k+p}$ поисковых испытаний следующей $(n+1)$-ой итерации определяются в соответствии с правилами.
-
-%Правила 1-4 параллельного алгоритма полностью повторяют правила 1-4 последовательного метода.
-
-%\textit{Правило 5}. Характеристики $R(i), 1 \leq i \leq k+1$, упорядочить в порядке убывания
-%\begin{equation}
-%R(t_1) \geq R(t_2) \geq \dots \geq R(t_{k-1}) \geq R(t_{k+1})
-%\end{equation}
-%и выбрать $p$ наибольших характеристик с номерами интервалов $t_j, 1 \leq j \leq p$.
-
-%\textit{Правило 6}. Провести новые испытания в точках $x^{k+j}, 1 \leq j \leq p$, вычисленных по формулам
-%\begin{equation}
-%x^{k+1}=\frac{x_{t_j}+x_{t_j-1}}{2},v(x_{t_j-1}) \neq v(x_{t_j}),
-%\end{equation}
-%\begin{equation}
-%x^{k+1}=\frac{x_{t_j}+x_{t_j-1}}{2} -sign(z_{t_j}-z_{t_j-1}) \frac{1}{2r_v} {\left[\frac{|z_{t_j}-z_{t_j-1}|}{\mu_v}\right]}^N, v=v(x_{t_j-1})=v(x_{t_j}).
-%\end{equation}
-
-%Алгоритм прекращает работу, если выполняется условие $\Delta_{t_j} \leq \varepsilon$ хотя бы для одного номера $t_j, 1 \leq j \leq p$, при этом одна из граничных точек, соответствующего этому номеру интервала, является внутренней точкой, т.е.
-$min{\{v_{t_j-1}, v_t_j\}} = 1$; здесь $\varepsilon>0$ есть заданная точность.
-
 \section{Программная реализация} \label{iOpt_discr}
@@
 Алгоритм глобального поиска, предназначенный для решения задач с не всюду вычислимой целевой функцией (АГП-Н), был реализован на базе фреймворка методов интеллектуальной оптимизации iOpt [\ref{rfa:rulit:iOptPaper}] с открытым исходным кодом [\ref{rfa:rulit:iOptGithub}].
+
+\begin{figure}[h!]
+	\center{\includegraphics[scale=0.82]{sheme.png}}
+	\caption{Блок-схема общего алгоритма функционирования фреймворка iOpt.}
+	\label{fig_iOpt}
+\end{figure}
 
@@
 Область применения фреймворка фокусируется на промышленных задачах и методах.
-В таких задачах часто возникает рассматриваемая в данном исследовании проблема частичной определенности критериев в отсутствии формульного описания исследуемой модели (модель вида «черный ящик»).
+В таких задачах часто возникает рассматриваемая в данном исследовании проблема частичной определенности критериев в отсутствии формульного описания исследуемой модели вида <<черный ящик>>.
 
@@
 
-\begin{figure}[h!]
-	\center{\includegraphics[scale=0.68]{figures/iOpt.pdf}}
-	\caption{Блок-схема общего алгоритма функционирования фреймворка iOpt}
-	\label{fig_iOpt}
-\end{figure}
-
 Реализация фреймворка iOpt выполнена с использованием объектно-ориентированного программирования в виде системы взаимосвязанных классов.
-Реализация описанного в разделе \ref{algorithm_discr} алгоритма содержится в базовом классе Method.
+Реализация описанного в разделе \ref{alg_discr} алгоритма содержится в базовом классе Method.
 Описанный далее в разделе \ref{GKLS_HC} генератор задач GKLS со скрытыми ограничениями (GKLS-HC) реализован в фреймворке iOpt в виде класса GKLSHiddenConstraint -- наследника интерфейса Problem.
@@
 
+\subsection{Решение серий тестовых задач} \label{GKLS_HC}
+
 Решение серии задач с известными глобальными оптимумами является одним из традиционных способов оценки качества работы методов глобального поиска. 
 Однако известные наборы тестовых задач не содержат областей невычислимости целевой функции, поэтому для тестирования разработанного алгоритма АГП-Н был предложен способ генерации тестовых задач со скрытыми ограничениями.
+
 В данном разделе дано краткое описание тестовых задач, а также приведены результаты экспериментов, демонстрирующие надёжность глобального поиска и эффективность разработанного алгоритма оптимизации.
 
-\subsection{Описание тестовых задач} \label{GKLS_HC}
+\subsubsection{Описание тестовых задач} \label{GKLS_HC}
 
@@
 
-\subsection{Решение тестовых задач с помощью АГП-Н}
+\subsubsection{Результаты решения тестовых задач с помощью АГП-Н}
 
@@
 Для полноценного сравнения алгоритм АГП-Н также был запущен на серии задач GKSL (т.е.
-на задачах без подобластей, в которых целевая функция является неопределенной).
+на задачах без подобластей, в которых целевая функция является неопределенной), что соответсвует поведению классического АГП.
+
+На Рис.
+\ref{oper_charact} показаны операционные характеристики, которые демонстрируют для каждого количества итераций число задач из тестовой выборки, решенное с заданной точностью $\varepsilon = 0.001$.
 Результаты эксперимента приведены в таблице \ref{iOpt_gkls}.
+
+Примеры визуализации процесса решения задачи GKLS-HC №~38 представлены на Рис.
+\ref{iOpt_result}.
+Черными точками отмечены точки испытаний, попавшие в подобласти неопределенности функции.
+Красной точкой отмечено найденное алгоритмом решение.
+
+\begin{figure}[h!]
+	\center{\includegraphics[scale=0.80]{figures/oper_charact_transp.pdf}}
+	\caption{Операционные характеристики алгоритмов АГП-Н и АГП}
+	\label{oper_charact}
+\end{figure}
+
+%\begin{table}[h!]
+%\centering
+%\caption{Результаты решения задач GKLS и GKLS-HC средствами фреймворка iOpt}
+%\begin{tabular}{||r|c|c|c||}
+%\hline
+%\multicolumn{1}{||c|}{} & alpha & \begin{tabular}[c]{@{}c@{}}среднее\\ число\\ итераций\end{tabular} & \begin{tabular}[c]{@{}c@{}}кол-во\%\ решенных\\ задач\end{tabular} \\ \hline
+%\multirow{2}{*}{GKLS-HC} & 0.08  & 2151 & 100 / 100 \\ \cline{2-4} 
+%                                        & 0.008 & 1635 & 100 / 100 \\ \hline
+%GKLS                                    & -     & 1510 & 100 / 100 \\ \hline
+%\end{tabular}
+%\label{iOpt_gkls}
+%\end{table}
 
@@
 \caption{Результаты решения задач GKLS и GKLS-HC средствами фреймворка iOpt}
-\begin{tabular}{||r|c|c|c||}
+\begin{tabular}{||r|c|c||}
 \hline
-\multicolumn{1}{||c|}{} & alpha & \begin{tabular}[c]{@{}c@{}}среднее\\ число\\ итераций\end{tabular} & \begin{tabular}[c]{@{}c@{}}кол-во\\ решенных\\ задач\end{tabular} \\ \hline
-\multirow{2}{*}{GKLS-HC} & 0.08  & 2151 & 100 / 100 \\ \cline{2-4} 
-                                        & 0.008 & 1635 & 100 / 100 \\ \hline
-GKLS                                    & -     & 1510 & 100 / 100 \\ \hline
+ класс задач & $\alpha$ & среднее число итераций \\ \hline
+\multirow{2}{*}{GKLS-HC} & 0.08 & 2151 \\ \cline{2-3} 
+ & 0.008 & 1635 \\ \hline
+GKLS & - & 1510 \\ \hline
 \end{tabular}
@@
 
-На Рис. \ref{oper_charact} показаны операционные характеристики, которые указывают для каждого количества итераций число задач из тестовой выборки, решенное с заданной точностью $\varepsilon = 0.001$.
-
-\begin{figure}[h!]
-	\center{\includegraphics[scale=0.75]{figures/oper_charact_transp.pdf}}
-	\caption{Операционные характеристики алгоритмов АГП-Н и АГП}
-	\label{oper_charact}
-\end{figure}
-
 Результаты проведенных экспериментов показывают, что после выполнения 2000 итераций АГП-Н с параметром $\alpha = 0.008$ будет получена оценка оптимума с требуемой точностью для 80\% задач.
@@
 
-Пример визуализации процесса решения задачи GKLS-HC №~38 представлен на Рис.
-\ref{iOpt_result}.
-Черными точками отмечены точки испытаний, попавшие в подобласти, в которых функция не определена.
-Красной точкой отмечено найденное алгоритмом решение.
-
 \begin{figure}[h!]
-	\center{\includegraphics[scale=1.1]{figures/iOpt_result.pdf}}
+	\center{\includegraphics[scale=0.93]{figures/iOpt_result.pdf}}
 	\caption{Графики линий уровня задачи GKLS-HC №~38 с точками проводимых испытаний, полученные средствами фреймворка iOpt: (а) $\alpha = 0.08$, (б) $\alpha = 0.008$ }
@@
 
+АГП-Н может быть использован не только в задаче настройки гиперараметров различных методов, но и для решения рядовых оптимизационных задач из различных прикладных областей.
+В связи с этим имеет смысл привести сравнение реализованного метода с существующими оптимизаторами на задачах с не всюду вычислимой целевой функцией.
+
 \subsubsection{Описание алгоритмов оптимизации}
@@
 
-В рамках экспериментов данного исследования были получены следующие результаты.
-
-При наличии у задачи скрытых ограничений худшее поведение демонстрируют алгоритмы differential\_evolution и brute.
-Алгоритм differential\_evolution полностью исчерпывает все выделенные ему итерации, не находя при этом корректный ответ.
+В рамках проведенного эксперимента были получены следующие результаты.
+
+При наличии у задачи скрытых ограничений худшее поведение демонстрируют алгоритмы \textit{differential\_evolution} и \textit{brute}.
+Алгоритм \textit{differential\_evolution} полностью исчерпывает все выделенные ему итерации, не находя при этом корректный ответ.
 Полный перебор в свою очередь в качестве результата работы может вернуть что угодно.
 
-Алгоритмы dual\_annealing и basinhopping в ряде случаев справляются с поставленной оптимизационной задачей.
-При этом в рамках эксперимента количество вычислений целевой функции алгоритмом dual\_annealing составило порядка 4000, в то время как basinhopping мог выполнять и 140000 вычислений для одной задачи.
-
-Алгоритмы direct и shgo демонстрируют хорошие результаты, хоть и не всегда обладают сходимостью.
-Среднее число итераций алгоритма direct составило 94 (соответствует 1981 вычислениям целевой функции).
-Алгоритм shgo в каждой задаче полностью исчерпал выделенные ему 200 итерации, выполняя в среднем 4482 вычислений целевой функции.
+Алгоритмы \textit{dual\_annealing} и \textit{basinhopping} в ряде случаев справляются с поставленной оптимизационной задачей.
+При этом в рамках эксперимента количество вычислений целевой функции алгоритмом \textit{dual\_annealing} составило порядка 4000, в то время как \textit{basinhopping} мог выполнять и 140000 вычислений для одной задачи.
+
+Алгоритмы \textit{direct} и \textit{shgo} демонстрируют хорошие результаты, хоть и не всегда обладают сходимостью.
+Среднее число итераций алгоритма \textit{direct} составило 94 (соответствует 1981 вычислениям целевой функции).
+Алгоритм \textit{shgo} в каждой задаче полностью исчерпал выделенные ему 200 итерации, выполняя в среднем 4482 вычислений целевой функции.
 В сравнении с данными методами АГП-Н работает эффективнее (отметим, что одна итерация АГП-Н соответствует одному вычислению значения целевой функции) и обладает лучшей сходимостью к глобальному оптимуму.
@@
 
-В таблице \ref{comparision} приведена информация о возможном поведении исследованных алгоритмов с задачами с не всюду определенной целевой функцией, обнаруженная при решении задач GKLS-HC.
+В таблице \ref{comparision} собрана информация о возможном поведении исследованных алгоритмов с задачами с не всюду определенной целевой функцией, обнаруженная при решении задач GKLS-HC.
 Алгоритм АГП-Н всегда находит глобальный минимум, тогда как differential\_evolution всегда возвращает значение nan.
@@
 
-\subsection{Решение задачи настройки гиперпараметров метода LinearSVC}
-
-В качестве модельной задачи оптимизации, в которой встречается недопустимое сочетание параметров, рассмотрим настройку гиперпараметров метода LinearSVC из библиотеки scikit-learn при решении задачи классификации.
+\subsection{Решение задач настройки методов машинного обучения}
+
+\subsubsection{Настройка гиперпараметров метода LinearSVC}
+
+В качестве первой модельной задачи оптимизации, в которой встречается недопустимое сочетание параметров, рассмотрим настройку гиперпараметров метода LinearSVC из библиотеки \textit{scikit-learn} при решении задачи классификации.
+
 Исследование проводилось на классическом датасете Iris, который содержит характеристики трех различных видов ириса и часто используется в качестве тестового примера для алгоритмов машинного обучения.
-Набор данных Iris состоит из 150 записей (по 50 записей на каждый вид ириса), содержащих по пять атрибутов.
-Датасет включен в библиотеку машинного обучения scikit-learn.
-
-Решалась задача максимизации метрики $Macro F1$ (характеризующей качество построения классификатора) в зависимости от трех гиперпараметров:
+Набор данных Iris состоит из 150 записей (по 50 записей на каждый вид ириса), содержащих по 5 атрибутов.
+Датасет включен в библиотеку машинного обучения \textit{scikit-learn}.
+
+Задача заключалась в максимизации метрики $F1$, характеризующей качество построения классификатора. Для LinearSVC варьировались 2 дискретных и 1 непрерывный параметры:
 \begin{itemize}
-\item вещественного коэффициента регуляризации $C$, $C \in [1,6]$;
-\item дискретного типа функции потерь $loss$, $loss \in \{hinge, squared\_hinge\}$;
-\item дискретного параметра $dual$, $dual \in \{true, false\}$, который позволяет выбирать между решением двойственной или основной задачи оптимизации.
+\item[--] тип функции потерь $loss$, $loss \in \{hinge, squared\_hinge\}$;
+\item[--] параметр $dual$, $dual \in \{true, false\}$, который позволяет выбирать между решением двойственной или основной задачи оптимизации;
+\item[--] коэффициент регуляризации $C$, $C \in [1,6]$.
 \end{itemize}
-
-Для возможности воспроизведения результатов параметр $random\_state$ был зафиксирован в значении 10.
-Остальные параметры метода LinearSVC рассматривались заданными по умолчанию.
-Максимизация метрики $F1$ проводилась с использованием нескольких алгоритмов:
+Для возможности воспроизведения результатов параметр $random\_state$ был зафиксирован в значении 10. Остальные параметры метода LinearSVC рассматривались заданными по умолчанию.
+
+Отметим, что сочетание гиперпараметров $loss = hingle$ и $dual = False$ приводит к <<падению>> метода LinearSVC (метод выбрасывает исключение и не выдает значения целевой метрики).
+
+Поставленная задача решалась в области изменения гиперпараметров с использованием следующих алгоритмов:
 \begin{itemize}
-\item GridSearchCV -- метод полного перебора по равномерной сетке в области изменения гиперпараметров;
-\item  RandomizedSearchCV -- метод случайного поиска в области изменения гиперпараметров;
-\item  АГП-Н, предложенный в данной статье; в алгоритме использовался параметр $r=3.5$ и было задано ограничение на число испытаний $K_{max}=100$.
+\item[--] АГП-Н (предложенный в данной статье) с использованием параметра надежности $r=3.5$ и ограничением на число испытаний $K_{max}=100$;
+\item[--] GridSearchCV (метод полного перебора по равномерной сетке);
+\item[--] RandomizedSearchCV (метод случайного поиска).
 \end{itemize}
 
-Отметим, что сочетание гиперпараметров $loss = hingle$ и $dual = False$ приводит к «падению» метода LinearSVC (метод выдает выбрасывает исключение и не выдает значения целевой метрики).
-В процессе своей работы оба алгоритма из scikit-learn решили задачу с выдачей предупреждения о возникновении недопустимого сочетания параметров; было найдено значение целевой метрики $F1 = 0.96$.
-Алгоритм АГП-Н также решил поставленную задачу и нашел лучшее значение $F1 = 0.973$ при гиперпараметрах $C=1.3125$, $loss=squared\_hinge$, $dual = False$.
+Полученные результаты представлены в таблице \ref{tab22}.
+
+\begin{table}[h!]
+\centering
+\caption{Результаты настройки гиперпараметров метода LinearSVC при решении задачи классификации на датасете Iris.}\label{tab22}
+\begin{tabular}{|r|c|c|}
+\hline
+метод & финальная F1 \\
+\hline
+АГП-Н & 0.973 \\
+\hline
+GridSearchCV & 0.96 \\
+\hline
+RandomizedSearchCV & 0.96 \\
+\hline
+\end{tabular}
+\end{table}
+
+Отметим, что в процессе своей работы оба алгоритма библиотеки \textit{scikit-learn} решили задачу с выдачей предупреждения о возникновении недопустимого сочетания параметров.
+Алгоритм АГП-Н нашел лучшее значение $F1 = 0.973$ при сочетании значений гиперпараметров $C=1.3125$, $loss=squared\_hinge$, $dual = False$.
+
+%%%%% new
+
+\subsubsection{Настройка гиперпараметров метода предсказания значений временного ряда }
+
+В качестве второй модельной задачи рассмотрим задачу настройки гиперпараметров методов предсказания значений временного ряда (time series forecasting).
+Предсказание значений производилось с использованием фреймворка FEDOT %\cite{FEDOT,nikitin2022automated}
+(фреймворк с открытым исходным кодом для задач автоматизированного моделирования и машинного обучения, AutoML).
+FEDOT поддерживает полный цикл решения задачи, включающий препроцессинг, выбор модели, настройку, кроссвалидацию и т.д.
+Фреймворк использует ML-модели, в основном из стандартных библиотек \textit{sklearn}, \textit{statsmodels} и \textit{keras}, а также интегрирован с фреймворком GOLEM.
+
+Исследование проводилось на датасете \textit{montly beer prodaction}, который содержит 476 записей валидационной выборки. %\cite{MonthlyBeerDataset,MonthlyBeerArticle}
+
+Построенный для решения поставленной задачи в FEDOT пайплайн включает в себя комбинацию из двух преобразований: lagged и cgru.
+Lagged-преобразование выполняется путем взятия значения переменной в предыдущий момент времени и включения его в модель в качестве признака в текущий момент времени.
+Для этого данные временного ряда сдвигаются на определенное количество шагов, которые называются лагом или временной задержкой.
+В качестве используемой нейронной сети выступает рекуррентная сеть (GRU), использующая свертку (CGRU).
+
+%%%% КАРТИНКА ПОД ВОПРОСОМ
+%\begin{figure}[h!]
+%\center
+%\includegraphics[width=0.6\textwidth]{fig2.png}
+%\caption{Визуализация пайплайна с простейшей схемой архитектуры сети CGRU.} \label{fig2}
+%\end{figure}
+
+У lagged-преобразования имеется один параметр window\_size, который изменяется от 5 до 250 с шагом 1.
+Поскольку этот параметр имеет большое количество значений, при оптимизации, будем рассматривать его как непрерывный (с округлением до единиц).
+У CGRU варьировались 4 дискретных и 2 непрерывных параметра:
+\begin{itemize}
+\item[--] cnn1\_kernel\_size $\in \{4, 5\}$,
+\item[--] cnn2\_kernel\_size $\in \{5, 6\}$, 
+\item[--] cnn1\_output\_size $\in \{32, 64\}$, 
+\item[--] cnn2\_output\_size $\in \{32, 64\}$;
+\item[--] hidden\_size $\in [20, 200]$, 
+\item[--] learning\_rate $\in [0.0005, 0.005]$.
+\end{itemize}
+Остальные параметры были зафиксированы в значениях по умолчанию: batch\_size = 64, num\_epochs = 50, optimizer = 'adamw', loss = 'mse'.
+
+Таким образом, в решатель iOpt передавалась задача с 3 непрерывными параметрами и 4 дискретными параметрами, каждый из которых принимает 2 значения (всего 16 комбинаций).
+
+Отметим, что при некоторых сочетаниях параметров window\_size, cnn1\_kernel\_size, cnn1\_output\_size, cnn2\_kernel\_size и cnn2\_output\_size невозможно вычислить значение критерия (метод возвращает бесконечное значение функции).
+Такие точки рассматривались как невычислимые.
+
+Для сравнения поставленная задача также была решена при помощи известного фреймворка оптимизации гиперпараметров Optuna%\cite{OptunaURL,optuna}
+. Оба решателя были интегрированы в FEDOT и задействованы в вычислительном эксперименте.
+
+Для решения поставленной задачи было выставлено ограничение на 1000 поисковых испытаний.
+В iOpt после 950 <<глобальных>> итераций глобального метода запускалось локальное уточнение с использованием метода Хука-Дживса.
+%%%%\cite{Himmelblau72}
+
+\begin{figure}[h!]
+\center \includegraphics[scale=0.94]{fig3.png}
+\caption{Визуализация предсказания и истинного значения временного ряда. } \label{fig3}
+\end{figure}
+
+\begin{table}[h!]
+\centering
+\caption{Результаты решения задачи с использованием iOpt и Optuna.}\label{tab1}
+\begin{tabular}{|l|c|c|}
+\hline
+& время решения, сек. & финальная MSE \\
+\hline
+iOpt & 6574.2 & 13.5 \\
+\hline
+Optuna &  9554.1 & 13.4 \\
+\hline
+\end{tabular}
+\end{table}
+
+Результаты проведенного эксперимента отражены в таблице~\ref{tab1}.
+Число точек, в которых не получилось вычислить критерий оптимизации, составило порядка 10\% от общего числа точек поисковых испытаний.
+Кривые, соответствующие полученному предсказанию, приведены на Рис.~\ref{fig3}.
+Темная линия соответствует истинным значениям временного ряда, голубая -- предсказанным значениям, полученным после использования Optuna, желтая линия -- предсказанным значениям, полученным после использования iOpt.
+
+\subsection{Результаты экспериментов на сурогатных моделях}
+
+Дополнительная серия экспериментов для сравнения iOpt, Optuna и HyperOpt была проведена с использованием фреймворка JAHS-Bench-201%, описанного в \cite{jahsbench}
+.
+JAHS-Bench предоставляет удобный доступ к набору задач для определения оптимального сочетания гиперпараметров и архитектуры нейронной сети.
+Данный инструмент для сравнения HPO-алгоритмов содержит обширный набор данных о качестве и производительности нейронных сетей, состоящий примерно из 161 миллиона точек данных, каждая из которых состоит из 20 показателей (метрик производительности).
+
+Особенностью JAHS-Bench является использование суррогатных моделей, подобранных на основе данных о производительности, охватывающих различные задачи оптимизации (одиночные/множественные цели), что предоставляет возможность варьировать архитектуру нейросети (выбирая один из имеющихся вариантов), а также ее гиперпараметры.
+При этом использование суррогатных моделей позволяет оценить качество работы обученного классификатора при конкретном наборе гиперпараметров за доли секунды вместо нескольких часов, которые были бы затрачены на обучение реальной нейросети.
+В качестве основы для суррогатных моделей в JAHS-Bench используются три популярных датасета, применяемых для задач классификации -- CIFAR-10, Colorectal-Histology и Fashion MNIST.
+
+В проведенных  экспериментах настройка гиперпараметров с использованием фреймворков iOpt, Optuna и HyperOpt проводилась на фиксированной архитектуре нейронной сети, предложенной в JAHS-Bench в качестве архитектуры по умолчанию.
+Для настройки были выбраны традиционно наиболее значимые (с точки зрения исследователя) гиперпараметры: категориальный параметр выбора функции активации (принимающий три возможных значения ReLU, Mish, Hardswish), а также непрерывные параметры Learning Rate and Weight Decay в диапазонах $[10^{-3},1]$ и $[10^{-5},10^{-2}]$ соответственно.
+
+В качестве целевой функции использовалась метрика точности валидации, вычисленная после 200 эпох обучения.
+Точность поиска в критерии остановки всех методов была установлена равной 0.01, а максимальное число испытаний -- равным 200.
+Для алгоритмов из Optuna и HyperOpt достигнутые показатели усреднялись по 10 запускам.
+Результаты сравнения представлены в таблице \ref{tab:met-surr}.
+Значения, соответствующие наилучшему результату в среднем, выделены жирным шрифтом, а значения, соответствующие наилучшему результату для наихудшего случая, выделены красным.
+
+\begin{table*}[h!]
+\centering
+\caption{Значения оптимизированных метрик для алгоритмов, усредненные по 10 запускам (суррогатные модели)}
+\label{tab:met-surr}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{ccccccc}
+\hline
+\textbf{}                      & \textbf{iOpt}   & \textbf{Optuna(random)} & \textbf{Optuna(tpe)}     & \textbf{Optuna(cmaes)}   & \textbf{Optuna(nsgaii)} & \textbf{Hyperopt(tpe)} \\ \hline
+\textbf{cifar10}               & {\color[HTML]{FE0000} \textbf{88.394} } & 88.144 $\pm$ 0.21\%         & 88.287 $\pm$ 0.21\%          & 88.218 $\pm$ 0.15\%          & 88.063 $\pm$ 0.27\%         & 88.316 $\pm$ 0.11\%        \\
+\textbf{fashion\_mnist}        & 94.797          & 94.677 $\pm$ 0.18\%         & 94.837 $\pm$ 0.13\%          & {\color[HTML]{FE0000} \textbf{94.952 $\pm$ 0.08\%} } & 94.631 $\pm$ 0.35\%         & 94.759 $\pm$ 0.19\%        \\
+\textbf{colorectal\_histology} & {\color[HTML]{FE0000} \textbf{94.447} } & 94.013 $\pm$ 0.33\%         & \textbf{94.447 $\pm$ 0.17\%} & 94.194 $\pm$ 1.18\%          & 94.150 $\pm$ 0.36\%         & 94.233 $\pm$ 0.17\%        \\ \hline
+\end{tabular}%
+}
+\end{table*}
+
+Результаты показывают превосходство iOpt на датасетах CIFAR-10 и Colorectal-Histology.
+При использовании набора данных Fashion MNIST фреймворк iOpt превосходит HyperOpt, но уступает Optuna.
+Обратите внимание, что при использовании суррогатных моделей время вычисления целевой функции занимает всего доли секунды, поэтому результаты, связанные со сравнением времени выполнения алгоритма, для этого эксперимента не приводятся.
+
+%%%%%% end new
 
@@
 
-В данной работе исследовалась проблема поиска глобального минимума вычислительно трудоемкой функции вида <<черный ящик>>.
+\colorbox{orange}{исправить в соответсвии с введением и экспериментами ->} В данной работе исследовалась проблема поиска глобального минимума вычислительно трудоемкой функции вида <<черный ящик>>.
 По сравнению с традиционной формулировкой, используемой в глобальной оптимизации, рассмотренная задача имеет важное отличие: предполагается, что целевая функция может быть не определена в некоторых подобластях области поиска (в частном случае, в одной или нескольких ее точках).
@@
 Разработанный в ходе исследования и реализованный в библиотеке iOpt метод АГП-Н успешно решил поставленную задачу, найдя лучшее сочетание гиперпараметров по сравнению со стандартными алгоритмами из библиотеки scikit-learn.
+\colorbox{orange}{<-}
 
@@
 
+\item \label{rfa:rulit:Zhou2021}
+Zhou~J.
+, Qiu~Y.,  Zhu~S., Armaghani~D.
+J.
+, Li~C., Nguyen~H.
+, Yagiz~S.
+Optimization of support vector machine through the use of metaheuristic algorithms in forecasting {TBM} advance rate~// Eng.
+Appl.
+Artif.
+Intell.
+2021.
+Vol.~97.
+p.~104015.
+DOI: 10.1016/j.engappai.2020.104015.
+
+\item \label{rfa:rulit:Yang2022}
+Yang~W., Xia~K., Fan~S., Wang~L., Li~T., Zhang~J., Feng~Y.
+A Multi-Strategy Whale Optimization Algorithm and Its Application~// Eng.
+Appl.
+Artif.
+Intell.
+2022.
+Vol.~108.
+p.~104558.
+DOI: 10.1016/j.engappai.2021.104558.
+
+\item \label{rfa:rulit:Frazier2018}
+Frazier~P.I. A Tutorial on Bayesian Optimization~// arXiv. 2018. DOI: 10.48550/arXiv.1807.02811.
+
+\item \label{rfa:rulit:Archetti2019}
+Archetti~F., Candelieri~A. Bayesian Optimization and Data Science. Cham: Springer Briefs in Optimization, 2019. DOI: 10.1007/978-3-030-24494-1.
+
 %9
 \item \label{rfa:rulit:Jones2021}
-Jones~D., Martins~J. The direct algorithm: 25 years later~// J. Glob. Optim. 2021. Vol. 79, No,~3. P. 521--566. DOI: 10.1007/s10898-020-00952-6.
+Jones~D., Martins~J. The direct algorithm: 25 years later~// J. Glob. Optim. 2021. Vol.~79, No.~3. P. 521--566. DOI: 10.1007/s10898-020-00952-6.
 
@@
 \item \label{rfa:rulit:Birect2020}
-Paulavi{\v c}ius~R., Sergeyev~Y.D., Kvasov~D.E., {\v Z}ilinskas~J. Globally-biased {BIRECT} algorithm with local accelerators for expensive global optimization~// 
-Expert Syst. Appl. 2020. Vol. 144. p. 113052. DOI: 10.1016/j.eswa.2019.113052.
+Paulavi{\v c}ius~R., Sergeyev~Y.D., Kvasov~D.E., {\v Z}ilinskas~J.
+Globally-biased {BIRECT} algorithm with local accelerators for expensive global optimization~// Expert Syst.
+Appl.
+2020.
+Vol.~144.
+p.~113052.
+DOI: 10.1016/j.eswa.2019.113052.
 
@@
 \item \label{rfa:rulit:Liberti2005}
-Liberti~L., Kucherenko~S. Comparison of deterministic and stochastic approaches to global optimization~// Int. Trans. Oper. Res. 2005. Vol. 12. P. 263--285.
+Liberti~L., Kucherenko~S. Comparison of deterministic and stochastic approaches to global optimization~// Int. Trans. Oper. Res. 2005. Vol.~12. P.~263--285.
 
@@
 \item \label{rfa:rulit:Sergeyev2018}
-Sergeyev~Y.D., Kvasov~D.E., Mukhametzhanov~M.S. On the efficiency of nature-inspired metaheuristics in expensive global optimization with limited budget~// Sci. Rep. 2018. Vol. 8, No.~1. p. 435.
+Sergeyev~Y.D., Kvasov~D.E., Mukhametzhanov~M.S. On the efficiency of nature-inspired metaheuristics in expensive global optimization with limited budget~// Sci. Rep. 2018. Vol.~8, No.~1. p.~435.
 
@@
 \item \label{rfa:rulit:Kvasov2013}
-Квасов~Д.Е., Сергеев~Я.Д. Методы липшицевой глобальной оптимизации в задачах управления~// Автоматика и телемеханика. 2013. № 9. C. 3--19.
+Квасов~Д.Е., Сергеев~Я.Д. Методы липшицевой глобальной оптимизации в задачах управления~// Автоматика и телемеханика. 2013. №~9. C.~3--19.
 
@@
 \item \label{rfa:rulit:Sergeyev2020}
-Sergeyev~Y.D., Candelieri~A., Kvasov~D.E., Perego~R. Safe global optimization of expensive noisy black-box functions in the $\delta$-Lipschitz framework~// 
-Soft Comput. 2020. Vol. 24, No.~23. P. 17715--17735. DOI: 10.1007/s00500-020-05030-3.
+Sergeyev~Y.D., Candelieri~A., Kvasov~D.E., Perego~R.
+Safe global optimization of expensive noisy black-box functions in the $\delta$-Lipschitz framework~// Soft Comput.
+2020.
+Vol.~24, No.~23.
+P.~17715--17735.
+DOI: 10.1007/s00500-020-05030-3.
 
@@
 \item \label{rfa:rulit:Dongarra2022}
-Dongarra~J.J. The evolution of mathematical software~// Commun. ACM. 2022. Vol. 65, No.~12. P. 66--72. DOI: 10.1145/3554977.
+Dongarra~J.J. The evolution of mathematical software~// Commun. ACM. 2022. Vol.~65, No.~12. P.~66--72. DOI: 10.1145/3554977.
 
@@
 \item \label{rfa:rulit:Duwe2020}
-Duwe~K., et al. State of the art and future trends in data reduction for high-performance computing~// Supercomput. Front. Innov. 2020. Vol. 7, No.~1. DOI: 10.14529/jsfi200101.
+Duwe~K., et al. State of the art and future trends in data reduction for high-performance computing~// Supercomput. Front. Innov. 2020. Vol.~7, No.~1. DOI: 10.14529/jsfi200101.
 
@@
 \item \label{rfa:rulit:Stripinis2021}
-Stripinis~L., Paulavi{\v c}ius~R. A new {DIRECT}-{GLh} algorithm for global optimization with hidden constraints~// Optim. Lett. 2021. Vol. 15, No.~6. P. 1865--1884.
+Stripinis~L., Paulavi{\v c}ius~R. A new {DIRECT}-{GLh} algorithm for global optimization with hidden constraints~// Optim. Lett. 2021. Vol.~15, No.~6. P.~1865--1884.
 DOI: 10.1007/s11590-021-01726-z.
@@
 \item \label{rfa:rulit:Audet2022}
-Audet~C., Batailly~A., Kojtych~S. Escaping unknown discontinuous regions in blackbox optimization~// SIAM J. Optim. 2022. Vol. 32, No.~3. P. 1843--1870. DOI: 10.1137/21m1420915.
+Audet~C., Batailly~A., Kojtych~S. Escaping unknown discontinuous regions in blackbox optimization~// SIAM J. Optim. 2022. Vol.~32, No.~3. P.~1843--1870. DOI: 10.1137/21m1420915.
 
@@
 Candelieri~A. Sequential model based optimization of partially defined functions
-under unknown constraints~// J. Glob. Optim. 2019. Vol. 79, No.~2. P. 281--303. DOI: 10.1007/s10898-019-00860-4.
+under unknown constraints~// J. Glob. Optim. 2019. Vol.~79, No.~2. P.~281--303. DOI: 10.1007/s10898-019-00860-4.
 
@@
 Strongin~R.G., Barkalov~K.A., Bevzuk~S.A. Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints~//
-Soft Comput. 2020. Vol. 24, No.~16. P. 11853--11865. DOI: 10.1007/s00500-020-05078-1.
+Soft Comput. 2020. Vol.~24, No.~16. P.~11853--11865. DOI: 10.1007/s00500-020-05078-1.
 
 \item \label{rfa:rulit:Usova2024}
-Usova~M.A., Barkalov~K.A.
-An Algorithm for Finding the Global Extremum of a Partially Defined Function~// Communications in Computer and Information Science.
-2024.
-Vol.
-1914.
-P.
-147--161.
-DOI: 10.1007/978-3-031-52470-7{\_}13.
+Usova~M.A., Barkalov~K.A. An Algorithm for Finding the Global Extremum of a Partially Defined Function~//
+Communications in Computer and Information Science. 2024. Vol.~1914. P.~147--161. DOI: 10.1007/978-3-031-52470-7{\_}13.
 
@@
 Barkalov~K.A., et al. On solving the problem of finding kinetic parameters of catalytic isomerization of the pentane-hexane fraction using a parallel global search algorithm~// 
-Mathematics. 2022. Vol. 10, No.~19. p. 3665. DOI: 10.3390/math10193665.
+Mathematics. 2022. Vol.~10, No.~19. p.~3665. DOI: 10.3390/math10193665.
 
@@
 Gubaydullin~I.M., Enikeeva~L.V., Barkalov~K.A., Lebedev~I.G., Silenko~D.G.
-Kinetic modeling of isobutane alkylation with mixed c4 olefins and sulfuric acid as a catalyst using the asynchronous global optimization algorithm~// Commun.
-Comput.
-Inf.
-Sci.
-2022.
-Vol.
-1618.
-P.
-293--306.
-DOI: 10.1007/978-3-031-11623-0{\_}20.
+Kinetic modeling of isobutane alkylation with mixed c4 olefins and sulfuric acid as a catalyst using the asynchronous global optimization algorithm~//
+Commun. Comput. Inf. Sci. 2022. Vol.~1618. P.~293--306. DOI: 10.1007/978-3-031-11623-0{\_}20.
 
@@
 \item \label{rfa:rulit:differential_evolution}
-Storn~R., Price~K., Differential Evolution - a Simple and Efficient Heuristic for Global Optimization over Continuous Spaces // Journal of Global Optimization. 1997. Vol. 11. P. 341-359.
+Storn~R., Price~K., Differential Evolution - a Simple and Efficient Heuristic for Global Optimization over Continuous Spaces //
+Journal of Global Optimization. 1997. Vol.~11. P.~341--359.
 
 \item \label{rfa:rulit:dual_annealing}
-Xiang~Y, Gubian~S, Suomela~B, Hoeng~J. Generalized Simulated Annealing for Efficient Global Optimization: the GenSA Package for R // The R Journal. 2013. Vol. 5, No. 1.
+Xiang~Y, Gubian~S, Suomela~B, Hoeng~J. Generalized Simulated Annealing for Efficient Global Optimization: the GenSA Package for R // The R Journal. 2013. Vol.~5, No.~1.
 
 \item \label{rfa:rulit:direct}
-Gablonsky~J., Kelley~C. A Locally-Biased form of the DIRECT Algorithm // Journal of Global Optimization. 2001. Vol. 21. P. 27-37.
+Gablonsky~J., Kelley~C. A Locally-Biased form of the DIRECT Algorithm // Journal of Global Optimization. 2001. Vol.~21. P.~27--37.
 
@@
 Parallel Global Search Algorithm with Local Tuning for Solving Mixed-Integer Global Optimization Problems // Lobachevskii Journal of Mathematics.
-Vol.
-7.
-No.
-42.
+Vol.~7, No.~42.
 2021.
-P.
-1492-1503.
+P.~1492--1503.
 
@@
 2023.
-С.
-179-185.
+С.~179--185.
 
@@
 2003.
-Vol.
-29, No.~4.
-P.
-469--480.
-
-\item \label{Archetti2019}
-Archetti F, Candelieri A (2019) Bayesian Optimization and Data Science. SpringerBriefs in Optimization, Springer, Cham%   \doi{10.1007/978-3-030-24494-1}
-
-\item \label{Frazier2018}
-Frazier PI (2018) A tutorial on bayesian optimization. arXiv %  \doi{10.48550/arXiv.1807.02811}
+Vol.~29, No.~4.
+P.~469--480.
 
 \end{enumerate}
+
 %\newline

### 2025-07-07 2025/Problems of Informatics/article.tex
Some minor corrections

@@
 Однако в большинстве методов ИИ и МО имеются гиперпараметры, от выбора конкретных значений которых может зависеть качество (в некоторой метрике) полученного решения.
-Тот факт, что разница в качестве получаемого решения при разных значениях гиперпараметров может быть достаточно значительной, привел к возникновению класса задач настройки гиперпараметров (HPO --- hyperparameter optimization).
+Тот факт, что разница в качестве получаемого решения при разных значениях гиперпараметров может быть достаточно значительной, привел к возникновению класса задач настройки гиперпараметров (HPO -- hyperparameter optimization).
 С математической точки зрения такие задачи соответствуют задачам глобальной оптимизации с фиксированными границами изменения переменных.
@@
 Модель итерационно обновляется на основе накапливаемой в процессе поиска оптимума информации, позволяя на каждой очередной итерации оценить наиболее вероятное положение глобального оптимума.
-Методы байесовской оптимизации обладают существенно  более  высокой  эффективностью  в  сравнении с полным перебором, но, тем не менее, также подвержены влиянию <<проклятия размерности>>, пусть и в меньшей степени.
+Методы байесовской оптимизации обладают более  высокой  эффективностью по сравнению с метаэвристическими алгоритмами, но в значительной степени подвержены влиянию <<проклятия размерности>>.
 
@@
 
-В рассматриваемых задачах настройки гиперпараметров актуальной становится новая проблема --- некорректная работа исследуемых алгоритмов ИИ и МО в некоторых (заранее не известных) подобластях области изменения гиперпараметров.
-При этом программные реализации настраиваемых методов могут вести себя абсолютно по-разному: в лучшем случае информируют пользователя о том, что задача не может быть решена, в худшем возвращают недопустимое значение (NaN, inf) в качестве найденного решения.
+В рассматриваемых задачах настройки гиперпараметров актуальной становится новая проблема -- некорректная работа исследуемых алгоритмов ИИ и МО в некоторых (заранее не известных) подобластях области изменения гиперпараметров.
+При этом программные реализации настраиваемых методов могут вести себя абсолютно по-разному: в лучшем случае информируют пользователя о том, что задача не может быть решена, в худшем -- возвращают недопустимое значение (NaN, inf) в качестве найденного решения.
 Проблема возникновения недопустимых комбинаций, которая не имеет особого значения в случае <<ручной>> настройки, стала камнем преткновения для многих фреймворков автоматической настройки гиперпараметров.
@@
 Продемонстрированы результаты численных экспериментов как с тестовыми задачами, так и с прикладной задачей настройки гиперпараметров метода LinearSVC при решении задачи классификации на известном наборе данных Iris.
+
+%TODO - поправить последнее предложение во введении, добавить туда фразу про новые эксперименты
 
@@
 Последнее предположение делает невозможными применение известного ин\-фор\-ма\-ци\-он\-но-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска [\ref{rfa:rulit:Strongin2000}] или других методов липшицевой оптимизации [\ref{rfa:rulit:PaulaviciusZilinskas2014},\ref{rfa:rulit:Sergeyev2017}].
-Для решения таких задач нами была разработана модификация алгоритма глобального поиска, основанная на одном из предложенных нами ранее подходов [\ref{rfa:rulit:Usova2024}].
+Для решения таких задач нами была разработана модификация алгоритма глобального поиска, основанная на предложенном ранее подходе из [\ref{rfa:rulit:Usova2024}].
 
@@
 
-Алгоритм глобального поиска предполагает построение последовательности точек $y^i$, так называемых, \textit{поисковых испытаний}, для которых вычисляются значения целевой функции $z^i = \varphi(y^i)$.
-Согласно используемой схеме редукции размерности, проведение испытания также требует вычисление значения $y^i=y(x^i)$.
+Алгоритм глобального поиска предполагает построение последовательности точек \textit{поисковых испытаний} $y^i \in D$, в которых вычисляются значения целевой функции $z^i = \varphi(y^i)$.
+Согласно используемой схеме редукции размерности, проведение испытания предполагает вычисление значения $y^i=y(x^i), x^i \in [0,1]$.
 При этом каждой точке испытания $x^i$ ставится в соответствие индекс $v^i$, определяемый по правилу
@@
 
-На каждом шаге алгоритма производится вычисление оценки константы Гёльдера, определение характеристик поисковых интервалов, выбор наиболее <<перспективного>> интервала и вычисление точки $y^{k+1}$ для проведения очередного испытания.
-
-Основные изменения в алгоритме затронули правило вычисления характеристики интервала (Алгоритм \ref{charact_alg}).
-Сходимость метода в этом случае требует доказательства, которое будет продемонстировано авторами в будущих статьях.
+На каждом шаге алгоритма производится вычисление оценки константы Гёльдера, определение характеристик поисковых интервалов, выбор наиболее перспективного интервала и вычисление точки $y^{k+1}$ для проведения очередного испытания.
+Правила вычисления характеристики интервала для задач с частично определенной целевой функцией представлены в виде Алгоритма \ref{charact_alg}. 
+%Сходимость метода в этом случае требует доказательства, которое будет продемонстировано авторами в будущих статьях.
 
@@
 
+%TODO - нужно ли здесь слово Алгоритм? Может быть, Вычисление характеристики поискового интервала?
+
 \begin{algorithm}
@@
 
-\textit{Примечание}. Здась и далее $\Delta_i$ используется для обозначаения длины $i$-го интервала в N-мерном пространстве, т.е.
+Значения $r$ и $\alpha$ являются параметрами алгоритма (их смысл пояснен ниже), а $\Delta_i$ обозначает длину $i$-го интервала в новой метрике,
 \[
@@
 
-Подробное описание правил алгоритма глобального поиска на случай не всюду вычислимой целевой функции (АГП-Н)  для решения задач (\ref{eq2}) приведено в Алгоритме \ref{main_alg}.
+Подробное описание правил алгоритма глобального поиска на случай не всюду вычислимой целевой функции (АГП-Н)  для решения задачи (\ref{eq2}) приведено в виде Алгоритма \ref{main_alg}.
 
@@
 
-Алгоритм глобального поиска, предназначенный для решения задач с не всюду вычислимой целевой функцией (АГП-Н), был реализован на базе фреймворка методов интеллектуальной оптимизации iOpt [\ref{rfa:rulit:iOptPaper}] с открытым исходным кодом [\ref{rfa:rulit:iOptGithub}].
+Алгоритм глобального поиска, предназначенный для решения задач с не всюду вычислимой целевой функцией (АГП-Н), был реализован на базе открытого фреймворка методов интеллектуальной оптимизации iOpt [\ref{rfa:rulit:iOptPaper},\ref{rfa:rulit:iOptGithub}]
+%с открытым исходным кодом [\ref{rfa:rulit:iOptGithub}].
 
@@
 Фреймворк позволяет проводить точную настройку параметров моделей и методов, используемых в прикладных исследованиях в различных научных областях.
-Характерными примерами задач, решаемых фреймворком iOpt, являются задачи настройки гиперпараметров методов машинного обучения, а также методов эвристической оптимизации.
-Область применения фреймворка фокусируется на промышленных задачах и методах.
+Характерными примерами задач, решаемых фреймворком iOpt, являются задачи настройки гиперпараметров методов машинного обучения.
 В таких задачах часто возникает рассматриваемая в данном исследовании проблема частичной определенности критериев в отсутствии формульного описания исследуемой модели вида <<черный ящик>>.
@@
 
-Реализация фреймворка iOpt выполнена с использованием объектно-ориентированного программирования в виде системы взаимосвязанных классов.
+Фреймворк iOpt написан на Python в виде системы взаимосвязанных классов.
 Реализация описанного в разделе \ref{alg_discr} алгоритма содержится в базовом классе Method.
-Описанный далее в разделе \ref{GKLS_HC} генератор задач GKLS со скрытыми ограничениями (GKLS-HC) реализован в фреймворке iOpt в виде класса GKLSHiddenConstraint -- наследника интерфейса Problem.
-В фреймворке добавлена возможность визуализации процесса решения задач со скрытыми ограничениями с использованием StaticPainterNDListener.
+Описанный далее в разделе \ref{GKLS_HC} генератор задач GKLS со скрытыми ограничениями (GKLS-HC) реализован в виде класса GKLSHiddenConstraint -- наследника интерфейса Problem.
+В фреймворк добавлена возможность визуализации процесса решения задач со скрытыми ограничениями с использованием StaticPainterNDListener.
 Доступны различные способы отрисовки линий уровня целевой функции: по равномерной сетке, только по точкам поисковых испытаний, по точкам испытаний с использованием интерполяции.
@@
 
-Решение серии задач с известными глобальными оптимумами является одним из традиционных способов оценки качества работы методов глобального поиска. 
-Однако известные наборы тестовых задач не содержат областей невычислимости целевой функции, поэтому для тестирования разработанного алгоритма АГП-Н был предложен способ генерации тестовых задач со скрытыми ограничениями.
+Решение серии задач с известными свойствами является одним из традиционных способов оценки качества работы методов глобального поиска. 
+Однако доступные наборы тестовых задач не содержат областей невычислимости целевой функции, поэтому для тестирования разработанного алгоритма АГП-Н был предложен способ генерации тестовых задач со скрытыми ограничениями.
 
@@
 
-Данные функции были дополнены областями невычислимости в форме эллипсоидов (координаты центров и длины радиусов генерировались случайно), при попадании в которые функция возвращает исключение, сообщающее о невозможности вычислить ее значение в заданной точке.
-
-В текущей версии фреймворк iOpt использует генератор GKLS для формирования задач класса Simple с размерностью от 2 до 5, числом локальных минимумов равным 10 и областью поиска от $-1$ до $1$ по каждой переменной.
+Данные функции были дополнены областями невычислимости в форме эллипсоидов (координаты центров и длины радиусов генерировались случайно), при попадании в которые функция возвращает исключение, сообщающее о невозможности вычислить ее значение.
+
+В текущей версии фреймворк iOpt использует генератор GKLS для формирования задач с размерностью от 2 до 5, числом локальных минимумов равным 10 и областью поиска от $-1$ до $1$ по каждой переменной.
 Радиус области притяжения глобального минимума и расстояние от глобального минимума до вершины квадратичной функции варьируется в зависимости от размерности задачи.
@@
 Для полноценного сравнения алгоритм АГП-Н также был запущен на серии задач GKSL (т.е.
-на задачах без подобластей, в которых целевая функция является неопределенной), что соответсвует поведению классического АГП.
+на задачах без подобластей, в которых целевая функция является неопределенной), что соответствует поведению исходного АГП.
 
@@
 
-SHGO (simplicial homology global optimization) сравнительно новый алгоритм глобальной оптимизации, основанный на применении симплициальной гомологии.
+SHGO (simplicial homology global optimization) -- сравнительно новый алгоритм глобальной оптимизации, основанный на применении симплициальной гомологии.
 Алгоритм использует концепции из комбинаторной интегральной теории гомологии для поиска подобластей, которые приблизительно локально выпуклы.
@@
 
-Алгоритмы \textit{differential\_evolution}, \textit{dual\_annealing}, \textit{basinhopping} относятся к классу стохастических алгоритмов (то есть работают с использованием случайных чисел и каждый запуск поиска может найти другое решение), потому поиск глобального минимума с использованием данных алгоритмов требует многократных запусков.
+Алгоритмы \textit{differential\_evolution}, \textit{dual\_annealing}, \textit{basinhopping} относятся к классу стохастических алгоритмов (т.е.
+каждый запуск поиска может найти другое решение), потому поиск глобального минимума с использованием данных алгоритмов требует многократных запусков.
 

### 2025-07-14 2025/Problems of Informatics/article.tex
added new conclusion and some fix by the text

@@
 \usepackage{xcolor}
+\usepackage{wasysym}
 
@@
 В статье изложено подробное описание модифицированного алгоритма глобального поиска для решения исследуемого класса задач.
-%Программная реализация алгоритма глобального поиска на случай невсюду вычислимой целевой функции (АГП-Н) была выполнена на базе фреймворка методов интеллектуальной эвристической оптимизации iOpt открытым исходным кодом.
+%Программная реализация алгоритма глобального поиска на случай не всюду вычислимой целевой функции (АГП-Н) была выполнена на базе фреймворка методов интеллектуальной эвристической оптимизации iOpt открытым исходным кодом.
 %Для проведения экспериментов был разработан генератор задач со скрытыми ограничениями GKLSHiddenConstraint, взявший за основу генератор GKLS, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами.
@@
 
-В рассматриваемых задачах настройки гиперпараметров актуальной становится новая проблема -- некорректная работа исследуемых алгоритмов ИИ и МО в некоторых (заранее не известных) подобластях области изменения гиперпараметров.
+В рассматриваемых задачах настройки гиперпараметров актуальной становится новая проблема -- некорректная работа исследуемых алгоритмов ИИ и МО в некоторых (заранее неизвестных) подобластях области изменения гиперпараметров.
 При этом программные реализации настраиваемых методов могут вести себя абсолютно по-разному: в лучшем случае информируют пользователя о том, что задача не может быть решена, в худшем -- возвращают недопустимое значение (NaN, inf) в качестве найденного решения.
 Проблема возникновения недопустимых комбинаций, которая не имеет особого значения в случае <<ручной>> настройки, стала камнем преткновения для многих фреймворков автоматической настройки гиперпараметров.
+
 Данное явление с математической точки зрения можно интерпретировать или как наличие в задаче некоторых скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], или как наличие неизвестных областей, в которых целевая функция не является непрерывной [\ref{rfa:rulit:Audet2022}], или как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Sergeyev2003},\ref{rfa:rulit:Strongin2020}].
@@
 Проведено исследование поведения других известных алгоритмов при решении задач с не всюду вычислимой целевой функцией.
-Продемонстрированы результаты численных экспериментов как с тестовыми задачами, так и с прикладной задачей настройки гиперпараметров метода LinearSVC при решении задачи классификации на известном наборе данных Iris.
-
-%TODO - поправить последнее предложение во введении, добавить туда фразу про новые эксперименты
+Продемонстрированы результаты численных экспериментов как с тестовыми задачами, так и несколькими прикладными задачами настройки гиперпараметров, в которых возникают недопустимые сочетания гиперпараметров.
+Выполнена настройка метода LinearSVC при решении задачи классификации на известном наборе данных Iris, а также метода предсказания значений временного ряда фреймворка FEDOT для датасета montly beer prodaction.
 
@@
 
-%TODO - нужно ли здесь слово Алгоритм? Может быть, Вычисление характеристики поискового интервала?
-
 \begin{algorithm}
-\caption{Алгоритм определения характеристики интервала}\label{charact_alg}
+\caption{Вычисление характеристики $i$-го интервала}\label{charact_alg}
 \begin{algorithmic}[1]
@@
             \State $ R \gets 2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}$
-\Else \Comment{ обе точки явлются невычислимыми или граничными }
+\Else \Comment{ обе точки являются невычислимыми или граничными }
            \State $R \gets \alpha{(1-\frac{1}{r})}^2 \Delta _i$
@@
 
-Подробное описание правил алгоритма глобального поиска на случай не всюду вычислимой целевой функции (АГП-Н)  для решения задачи (\ref{eq2}) приведено в виде Алгоритма \ref{main_alg}.
+Подробное описание правил алгоритма глобального поиска на случай не всюду вычислимой целевой функции для решения задачи (\ref{eq2}) приведено в виде Алгоритма \ref{main_alg}.
 
@@
 	\center{\includegraphics[scale=0.15]{figures/task_61.png}}
-	\caption{Распределение точек испытаний с различными видами невычислимых областей и настройкой параметра $\alpha$: (а) область невычислимости на границе области поиска, $\alpha = 1.0$; (б) случайно расположенные области невычислимости, $\alpha = 1.0$; (в) область невычислимости на границе области поиска, $\alpha = 0.08$; (г) случайно расположенные области невычислимости, $\alpha = 0.08$.}
+	\caption{Распределение точек испытаний с различными видами невычислимых областей и настройкой параметра $\alpha$: (а) область невычислимости на границе области поиска, $\alpha = 1.0$; (б) случайно расположенные области невычислимости, $\alpha = 1.0$; (в) область невычислимости на границе области поиска, $\alpha = 0.08$; (г) случайно расположенные области невычислимости, $\alpha = 0.08$}
 	\label{task_61}
@@
 	\center{\includegraphics[scale=0.82]{sheme.png}}
-	\caption{Блок-схема общего алгоритма функционирования фреймворка iOpt.}
+	\caption{Блок-схема общего алгоритма функционирования фреймворка iOpt}
 	\label{fig_iOpt}
@@
 
+Результаты эксперимента приведены в таблице \ref{iOpt_gkls}.
+
 На Рис.
 \ref{oper_charact} показаны операционные характеристики, которые демонстрируют для каждого количества итераций число задач из тестовой выборки, решенное с заданной точностью $\varepsilon = 0.001$.
-Результаты эксперимента приведены в таблице \ref{iOpt_gkls}.
-
-Примеры визуализации процесса решения задачи GKLS-HC №~38 представлены на Рис.
-\ref{iOpt_result}.
-Черными точками отмечены точки испытаний, попавшие в подобласти неопределенности функции.
-Красной точкой отмечено найденное алгоритмом решение.
+
+\begin{table}[h!]
+\centering
+\caption{Результаты решения задач GKLS и GKLS-HC средствами фреймворка iOpt}
+\begin{tabular}{|r|c|c|}
+\hline
+ класс задач & $\alpha$ & среднее число итераций \\ \hline\hline
+\multirow{2}{*}{GKLS-HC} & 0.08 & 2151 \\ \cline{2-3} 
+ & 0.008 & 1635 \\ \hline
+GKLS & - & 1510 \\ \hline
+\end{tabular}
+\label{iOpt_gkls}
+\end{table}
 
 \begin{figure}[h!]
-	\center{\includegraphics[scale=0.80]{figures/oper_charact_transp.pdf}}
+	\center{\includegraphics[scale=0.9]{figures/oper_charact_transp.pdf}}
 	\caption{Операционные характеристики алгоритмов АГП-Н и АГП}
@@
 
-\begin{table}[h!]
-\centering
-\caption{Результаты решения задач GKLS и GKLS-HC средствами фреймворка iOpt}
-\begin{tabular}{||r|c|c||}
-\hline
- класс задач & $\alpha$ & среднее число итераций \\ \hline
-\multirow{2}{*}{GKLS-HC} & 0.08 & 2151 \\ \cline{2-3} 
- & 0.008 & 1635 \\ \hline
-GKLS & - & 1510 \\ \hline
-\end{tabular}
-\label{iOpt_gkls}
-\end{table}
+Примеры визуализации процесса решения задачи GKLS-HC №~38 представлены на Рис.
+\ref{iOpt_result}.
+Черными точками отмечены точки испытаний, попавшие в подобласти неопределенности функции.
+Красной точкой отмечено найденное алгоритмом решение.
+
+\begin{figure}[h!]
+	\center{\includegraphics[scale=1]{figures/iOpt_result.pdf}}
+	\caption{Графики линий уровня задачи GKLS-HC №~38 с точками проводимых испытаний, полученные средствами фреймворка iOpt: (а) $\alpha = 0.08$, (б) $\alpha = 0.008$ }
+	\label{iOpt_result}
+\end{figure}
 
@@
 Отметим, что за счёт регулировки параметра $\alpha$ получилось добиться такой же скорости работы, как у  АГП при решении серии задач GKLS (в терминах числа испытаний, требующихся для решения задач с заданной точностью).
-
-\begin{figure}[h!]
-	\center{\includegraphics[scale=0.93]{figures/iOpt_result.pdf}}
-	\caption{Графики линий уровня задачи GKLS-HC №~38 с точками проводимых испытаний, полученные средствами фреймворка iOpt: (а) $\alpha = 0.08$, (б) $\alpha = 0.008$ }
-	\label{iOpt_result}
-\end{figure}
 
@@
 \caption{Возможные выявленные результаты работы алгоритмов глобальной оптимизации библиотеки \textit{scipy.optimize} на задачах GKLS-HC}
-\begin{tabular}{||r|c|c|c|c||}
+\begin{tabular}{|r|c|c|c|c|}
 \hline
-\multirow{2}{*}{алгоритм} & \multicolumn{4}{c||}{возможный результат поиска} \\ \cline{2-5} 
- & \multicolumn{1}{c|}{глобальный минимум} & \multicolumn{1}{c|}{локальный минимум} & \multicolumn{1}{c|}{nan} & \multicolumn{1}{c||}{1e+100} \\ \hline
-АГП-Н                   &     \checkmark     &                   &              &             \\ \hline
-differential\_evolution &                    &                   &  \checkmark  &             \\ \hline
-dual\_annealing         &     \checkmark     &                   &  \checkmark  &             \\ \hline
-basinhopping            &     \checkmark     &    \checkmark     &              &  \checkmark \\ \hline
-direct                  &     \checkmark     &    \checkmark     &              &             \\ \hline
-shgo                    &     \checkmark     &    \checkmark     &              &             \\ \hline
-brute                   &     \checkmark     &    \checkmark     &  \checkmark  &  \checkmark \\ \hline
+\multirow{2}{*}{алгоритм} & \multicolumn{4}{c|}{возможный результат поиска} \\ \cline{2-5} 
+ & \multicolumn{1}{c|}{глобальный минимум} & \multicolumn{1}{c|}{локальный минимум} & \multicolumn{1}{c|}{nan} & \multicolumn{1}{c|}{1e+100} \\ \hline\hline
+АГП-Н                   &     \CIRCLE   &                   &              &             \\ \hline
+differential\_evolution &                    &                   &  \CIRCLE  &             \\ \hline
+dual\_annealing         &     \CIRCLE     &                   &  \CIRCLE  &             \\ \hline
+basinhopping            &     \CIRCLE     &    \CIRCLE    &              &  \CIRCLE \\ \hline
+direct                  &     \CIRCLE    &    \CIRCLE     &              &             \\ \hline
+shgo                    &     \CIRCLE     &    \CIRCLE     &              &             \\ \hline
+brute                   &     \CIRCLE     &    \CIRCLE    &  \CIRCLE  &  \CIRCLE \\ \hline
 \end{tabular}
@@
 Полученные результаты представлены в таблице \ref{tab22}.
+Отметим, что в процессе своей работы оба алгоритма библиотеки \textit{scikit-learn} решили задачу с выдачей предупреждения о возникновении недопустимого сочетания параметров.
+Алгоритм АГП-Н нашел лучшее значение $F1 = 0.973$ при сочетании значений гиперпараметров $C=1.3125$, $loss=squared\_hinge$, $dual = False$.
 
@@
 \centering
-\caption{Результаты настройки гиперпараметров метода LinearSVC при решении задачи классификации на датасете Iris.}\label{tab22}
+\caption{Результаты настройки гиперпараметров метода LinearSVC при решении задачи классификации на датасете Iris}\label{tab22}
 \begin{tabular}{|r|c|c|}
@@
 метод & финальная F1 \\
-\hline
+\hline \hline
 АГП-Н & 0.973 \\
@@
 
-Отметим, что в процессе своей работы оба алгоритма библиотеки \textit{scikit-learn} решили задачу с выдачей предупреждения о возникновении недопустимого сочетания параметров.
-Алгоритм АГП-Н нашел лучшее значение $F1 = 0.973$ при сочетании значений гиперпараметров $C=1.3125$, $loss=squared\_hinge$, $dual = False$.
-
 %%%%% new
@@
 FEDOT поддерживает полный цикл решения задачи, включающий препроцессинг, выбор модели, настройку, кроссвалидацию и т.д.
-Фреймворк использует ML-модели, в основном из стандартных библиотек \textit{sklearn}, \textit{statsmodels} и \textit{keras}, а также интегрирован с фреймворком GOLEM.
+Фреймворк использует ML-модели, в основном из стандартных библиотек \textit{sklearn}, \textit{statsmodels} и \textit{keras}.
 
@@
 %\center
-%\includegraphics[width=0.6\textwidth]{fig2.png}
+%\includegraphics[width=0.4\textwidth]{fig2.png}
 %\caption{Визуализация пайплайна с простейшей схемой архитектуры сети CGRU.} \label{fig2}
@@
 
-Таким образом, в решатель iOpt передавалась задача с 3 непрерывными параметрами и 4 дискретными параметрами, каждый из которых принимает 2 значения (всего 16 комбинаций).
-
+Таким образом, фреймворк iOpt решал задачу минимизации метрики $MSE$, варьируя 3 непрерывных и 4 дискретных параметра, каждый из которых принимал 2 значения (всего 16 комбинаций).
 Отметим, что при некоторых сочетаниях параметров window\_size, cnn1\_kernel\_size, cnn1\_output\_size, cnn2\_kernel\_size и cnn2\_output\_size невозможно вычислить значение критерия (метод возвращает бесконечное значение функции).
@@
 %%%%\cite{Himmelblau72}
-
-\begin{figure}[h!]
-\center \includegraphics[scale=0.94]{fig3.png}
-\caption{Визуализация предсказания и истинного значения временного ряда. } \label{fig3}
-\end{figure}
+Результаты проведенного эксперимента отражены в таблице~\ref{tab1}.
 
@@
 \centering
-\caption{Результаты решения задачи с использованием iOpt и Optuna.}\label{tab1}
+\caption{Результаты решения задачи настройки гиперпараметров метода предсказания значений временного ряда на датасете montly beer prodaction с использованием фреймворков iOpt и Optuna}\label{tab1}
 \begin{tabular}{|l|c|c|}
 \hline
-& время решения, сек. & финальная MSE \\
-\hline
+фреймворк & время решения, сек. & финальная MSE \\
+\hline \hline
 iOpt & 6574.2 & 13.5 \\
@@
 
-Результаты проведенного эксперимента отражены в таблице~\ref{tab1}.
 Число точек, в которых не получилось вычислить критерий оптимизации, составило порядка 10\% от общего числа точек поисковых испытаний.
@@
 
-\subsection{Результаты экспериментов на сурогатных моделях}
-
-Дополнительная серия экспериментов для сравнения iOpt, Optuna и HyperOpt была проведена с использованием фреймворка JAHS-Bench-201%, описанного в \cite{jahsbench}
-.
-JAHS-Bench предоставляет удобный доступ к набору задач для определения оптимального сочетания гиперпараметров и архитектуры нейронной сети.
-Данный инструмент для сравнения HPO-алгоритмов содержит обширный набор данных о качестве и производительности нейронных сетей, состоящий примерно из 161 миллиона точек данных, каждая из которых состоит из 20 показателей (метрик производительности).
-
-Особенностью JAHS-Bench является использование суррогатных моделей, подобранных на основе данных о производительности, охватывающих различные задачи оптимизации (одиночные/множественные цели), что предоставляет возможность варьировать архитектуру нейросети (выбирая один из имеющихся вариантов), а также ее гиперпараметры.
-При этом использование суррогатных моделей позволяет оценить качество работы обученного классификатора при конкретном наборе гиперпараметров за доли секунды вместо нескольких часов, которые были бы затрачены на обучение реальной нейросети.
-В качестве основы для суррогатных моделей в JAHS-Bench используются три популярных датасета, применяемых для задач классификации -- CIFAR-10, Colorectal-Histology и Fashion MNIST.
-
-В проведенных  экспериментах настройка гиперпараметров с использованием фреймворков iOpt, Optuna и HyperOpt проводилась на фиксированной архитектуре нейронной сети, предложенной в JAHS-Bench в качестве архитектуры по умолчанию.
-Для настройки были выбраны традиционно наиболее значимые (с точки зрения исследователя) гиперпараметры: категориальный параметр выбора функции активации (принимающий три возможных значения ReLU, Mish, Hardswish), а также непрерывные параметры Learning Rate and Weight Decay в диапазонах $[10^{-3},1]$ и $[10^{-5},10^{-2}]$ соответственно.
-
-В качестве целевой функции использовалась метрика точности валидации, вычисленная после 200 эпох обучения.
-Точность поиска в критерии остановки всех методов была установлена равной 0.01, а максимальное число испытаний -- равным 200.
-Для алгоритмов из Optuna и HyperOpt достигнутые показатели усреднялись по 10 запускам.
-Результаты сравнения представлены в таблице \ref{tab:met-surr}.
-Значения, соответствующие наилучшему результату в среднем, выделены жирным шрифтом, а значения, соответствующие наилучшему результату для наихудшего случая, выделены красным.
-
-\begin{table*}[h!]
-\centering
-\caption{Значения оптимизированных метрик для алгоритмов, усредненные по 10 запускам (суррогатные модели)}
-\label{tab:met-surr}
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{ccccccc}
-\hline
-\textbf{}                      & \textbf{iOpt}   & \textbf{Optuna(random)} & \textbf{Optuna(tpe)}     & \textbf{Optuna(cmaes)}   & \textbf{Optuna(nsgaii)} & \textbf{Hyperopt(tpe)} \\ \hline
-\textbf{cifar10}               & {\color[HTML]{FE0000} \textbf{88.394} } & 88.144 $\pm$ 0.21\%         & 88.287 $\pm$ 0.21\%          & 88.218 $\pm$ 0.15\%          & 88.063 $\pm$ 0.27\%         & 88.316 $\pm$ 0.11\%        \\
-\textbf{fashion\_mnist}        & 94.797          & 94.677 $\pm$ 0.18\%         & 94.837 $\pm$ 0.13\%          & {\color[HTML]{FE0000} \textbf{94.952 $\pm$ 0.08\%} } & 94.631 $\pm$ 0.35\%         & 94.759 $\pm$ 0.19\%        \\
-\textbf{colorectal\_histology} & {\color[HTML]{FE0000} \textbf{94.447} } & 94.013 $\pm$ 0.33\%         & \textbf{94.447 $\pm$ 0.17\%} & 94.194 $\pm$ 1.18\%          & 94.150 $\pm$ 0.36\%         & 94.233 $\pm$ 0.17\%        \\ \hline
-\end{tabular}%
-}
-\end{table*}
-
-Результаты показывают превосходство iOpt на датасетах CIFAR-10 и Colorectal-Histology.
-При использовании набора данных Fashion MNIST фреймворк iOpt превосходит HyperOpt, но уступает Optuna.
-Обратите внимание, что при использовании суррогатных моделей время вычисления целевой функции занимает всего доли секунды, поэтому результаты, связанные со сравнением времени выполнения алгоритма, для этого эксперимента не приводятся.
+\begin{figure}[h!]
+\center \includegraphics[scale=1.1]{fig3.png}
+\caption{Визуализация предсказания и истинного значения временного ряда} \label{fig3}
+\end{figure}
 
@@
 
-\colorbox{orange}{исправить в соответсвии с введением и экспериментами ->} В данной работе исследовалась проблема поиска глобального минимума вычислительно трудоемкой функции вида <<черный ящик>>.
-По сравнению с традиционной формулировкой, используемой в глобальной оптимизации, рассмотренная задача имеет важное отличие: предполагается, что целевая функция может быть не определена в некоторых подобластях области поиска (в частном случае, в одной или нескольких ее точках).
-В таких подобластях невозможно корректно провести численное моделирование и вычислить значение целевой функции.
-
-Основная трудность рассматриваемых задач заключается в невозможности заранее узнать о подобластях невычислимости; фактически, область допустимых сочетаний параметров является заранее неопределенной.
-Данное явление можно интерпретировать как наличие в задаче некоторых скрытых ограничений.
-
-В статье изложено подробное описание модифицированного алгоритма глобального поиска АГП-Н для решения исследуемого класса задач и генератора тестовых функций GKLS-HC, основанного на известном генераторе GKLS.
-Приведенные в статье результаты экспериментов, полученные при решении серии тестовых задач, демонстрируют надёжность разработанного алгоритма.
-Результаты работы АГП-Н сравнивались с результатами, полученными с помощью стандартных алгоритмов оптимизации из библиотеки \textit{scipy.optimize}.
-В рамках данного эксперимента разработанный алгоритм показал свою эффективность.
-
-Также с помощью АГП-Н была решена задача настройки гиперпараметров алгоритма классификации LinearSVC.
-Известно, что сочетание некоторых гиперпараметров LinearSVC приводит к <<падению>> метода.
-Разработанный в ходе исследования и реализованный в библиотеке iOpt метод АГП-Н успешно решил поставленную задачу, найдя лучшее сочетание гиперпараметров по сравнению со стандартными алгоритмами из библиотеки scikit-learn.
-\colorbox{orange}{<-}
+В настоящей работе были рассмотрены актуальные проблемы решения задач настройки гиперпараметров методов искусственного интеллекта (ИИ) и машинного обучения (МО).
+
+Основная сложность таких задач заключается в довольно длительном решении соответствующей исходной задачи для каждого выбранного сочетания гиперпараметров, что накладывает ограничение на возможное количество испытываемых комбинаций.
+
+В свою очередь принципиально новой и не менее сложной проблемой является некорректная работа исследуемых алгоритмов ИИ и МО в некоторых заранее неизвестных подобластях области изменения гиперпараметров; фактически, область допустимых сочетаний параметров является заранее неопределенной.
+Проблема возникновения таких недопустимых комбинаций, которая ранее не имела особого значения при <<ручной>> настройке, стала камнем преткновения для многих фреймворков автоматической настройки гиперпараметров.
+
+Для решения исследуемого класса задач была разработана модификация одного из эффективных детерминированных методов решения задач липшицевой глобальной оптимизации -- ин\-фор\-ма\-ци\-он\-но-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска.
+Разработанный алгоритм реализован на базе фреймворка методов интеллектуальной оптимизации с открытым исходным кодом iOpt.
+
+В статье изложено подробное описание модифицированного алгоритма (АГП-Н).
+Приведены результаты серийных экспериментов, демонстрирующие надёжность разработанного подхода.
+Продемонстрировано сравнение результатов работы АГП-Н с результатами, полученными с помощью стандартных алгоритмов оптимизации из библиотеки SciPy.
+Показана эффективность разработанного метода при решении исследуемого класса задач.
+
+В качестве демонстрационных примеров были решены несколько задач настройки гиперпараметров, в которых возникает проблема недопустимых сочетаний гиперпараметров, приводящих к <<падению>> метода.
+
+В качестве первой модельной задачи была рассмотрена задача настройки гиперпараметров классического алгоритма классификации LinearSVC.
+Разработанный метод успешно решил поставленную задачу, найдя лучшее сочетание гиперпараметров по сравнению со стандартными алгоритмами из библиотеки scikit-learn.
+
+В качестве второй модельной задачи была рассмотрена задача предсказания значений временного ряда средствами фреймворка FEDOT.
+В ходе эксперимента также было проведено сравнение с другим известным фреймворком, применяемым для настройки гиперпараметров -- Optuna.
+В ходе эксперимента фреймворк iOpt продемонстрировал свое преимущество во времени решения задачи, немного уступив в качестве.
 
@@
 , Yagiz~S.
-Optimization of support vector machine through the use of metaheuristic algorithms in forecasting {TBM} advance rate~// Eng.
-Appl.
-Artif.
-Intell.
+Optimization of support vector machine through the use of metaheuristic algorithms in forecasting {TBM} advance rate~// Engineering Applications of Artificial Intelligence.
 2021.
@@
 Н.И.
-Лобачевского, e-mail:anton.shtanyuk@itmm.unn.ru.
+Лобачевского.
+E-mail: anton.shtanyuk@itmm.unn.ru.
 Количество печатных работ 53.
@@
 He is the author or coauthor more than 30 papers in these areas.
-\paragraph{Anton Shtanyuk} -- PhD., Associate Professor at the Department of Informatics and Research Automation of the Institute of Information Technology, mathematics and mechanics at Lobachevsky State University of Nizhny Novgorod, mail:anton.shtanyuk@itmm.unn.ru The number of publications: 50.
+\paragraph{Anton Shtanyuk} -- PhD., Associate Professor at the Department of Informatics and Research Automation of the Institute of Information Technology, mathematics and mechanics at Lobachevsky State University of Nizhny Novgorod.
+E-mail: anton.shtanyuk@itmm.unn.ru The number of publications: 50.
 Research interests: modelling and design of complex systems, programming languages and technology, programming teaching, algorithms and data structures.

### 2025-07-20 2025/MMST_DecisionTree/ParallelDecisionTree.tex
Статья о синхронном параллельном запуске локальных методов Начальная версия

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%

\usepackage[utf8]{inputenc} % кодировка
\usepackage[english, russian]{babel} % Русские и английские переносы
\usepackage{cite}              % для корректного оформления литературы
\usepackage{enumitem}
\usepackage{amsmath}

\begin{document}
%
\title{Параллельный запуск локальных методов алгоритме глобального поиска с использованием деревьев решений}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{К.А. Баркалов\inst{1}\orcidID{0000-1111-2222-3333} \and
И.Г. Лебедев\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Я.В. Селенко\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Нижегородский государственный университет им. Н.И. Лобачевского, Нижний Новгород, Российская Федерация }
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
В данной работе предлагается параллельный подход к запуску локальных методов при решении задачи глобальной оптимизации алгоритмом глобального поиска. Целевая функция задачи задана как черный ящики и от нее требуется лишь удовлетворять условию Липшица с априори неизвестной константой. В работе рассматривается метод выбора окрестности локальных экстремумов целевой функции на основе анализа накопленной поисковой информации. Проведение такого анализа с использованием методов машинного обучения позволяет принять решение о параллельном запуске локальных методов, что может ускорить сходимость алгоритма. Это предположение было подтверждено результатами численных экспериментов, демонстрирующих ускорение при решении ряда тестовых задач.

\keywords{
	Глобальная оптимизация $\cdot$ 
	Многоэкстремальные функции $\cdot$ 
	Параллельные вычисления $\cdot$ 
	Машинное обучение $\cdot$ 
	Древо решений $\cdot$}
\end{abstract}
%
%
%

\section{Введение}

В настоящей работе рассматриваются параллельные алгоритмы решения многомерных задач глобальной оптимизации. Такие задачи часто возникают в тех случаях, когда необходимо подобрать значения параметров исследуемой математической модели, при которых результаты моделирования лучше всего соответствуют экспериментальным данным. Для решения задач этого класса известно множество алгоритмов: от метаэвристических, основанных на идее случайного поиска \cite{Ferreiro2013,Garcia2014,Langdon2011}, до детерминированных алгоритмов, гарантирующих сходимость к глобальному минимуму \cite{Evtushenko2009,He2008,Paulavicius2011}.

Поскольку в реальных задачах глобальной оптимизации каждое вычисление значения функции (далее \textit{испытание}) является операцией, требующей больших вычислительных затрат, количество таких попыток приходится сокращать. Это может быть достигнуто за счет преднамеренного выбора вариантов в ходе поиска оптимального решения, отсекая бесперспективные поисковые под области и исследуя только те, в которых можно найти решение задачи. Алгоритм глобального поиска (АГП) основан на этой идее \cite{Strongin2000}. В настоящей работе мы попытались объединить АГП и метод локальной оптимизации \cite{Nocedal, Kelley}  (метод поиска по образцу Хука-Дживса \cite{HookJeeves}), чтобы уменьшить количество выполняемых испытаний. Решение о запуске локального метода будет приниматься с использованием дерева решений.

\section{Алгоритмы}\label{SecA}

Рассмотрим задачу поиска глобального минимума функции $\varphi(y)$ на гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1\le\ i\le\ n \}$. 
\begin{equation} \label{sec:problem}   
	\varphi(y^*) = min\{\varphi(y):y\in D\}, D = \{y \in R^N : a_i \leq y_i \leq b_i, 1 \leq i \leq N \},
\end{equation}
где $a,b \in R$ — заданные векторы.

Также предположим, что функция удовлетворяет условию Липшица с \textit{априори} неизвестной константой $L$, что соответствует ограниченному изменению значений функции при ограниченном изменении аргумента.
\begin{displaymath} 
	|\varphi(y_1)-\varphi(y_2)|\leq L\parallel y_1-y_2 \parallel ,y_1,y_2 \in D, 0<L< \infty. 
\end{displaymath}
Это предположение можно рассматривать (применительно к реальным задачам) как отражение ограниченной мощности, порождающей изменения в моделируемой системе.

Численное решение задачи (\ref{sec:problem}) сводится к построению оценки $y_k^\ast\in D$, соответствующей некоторому понятию близости к точке $y^\ast$ (например, ${ ||y^\ast-y}_k^\ast||\le\ \varepsilon$, где $\varepsilon\geq0$ — заданная точность) на основе конечного числа $k$ вычисленных значений целевой функции. Применительно к классу рассматриваемых задач предполагается, что целевая функция $\varphi(y)$ может быть определена алгоритмически, в результате выполнения некоторой подпрограммы или библиотеки.

При решении многомерных задач применяется уменьшение размерности (т.е. сведение многомерной задачи к эквивалентной одномерной) с помощью кривых Пеано. Они позволяют свести многомерную задачу оптимизации в области $D$ к одномерной задаче минимизации в интервале $[0, 1]$.
\begin{displaymath}
	\varphi(y(x^\ast))\ =\min\{\varphi(y(x)): x\in [0,1]\},
\end{displaymath}
где функция $\varphi(y(x^\ast))$ удовлетворяет более общему условию Гельдера
\begin{displaymath}
	\left|\varphi (y \left(x_1\right))- \varphi (y \left(x_2\right)\right )|\le\ H\left|x_1-x_2\right|^\frac{1}{N},\ x_1,\ x_2\epsilon[0,1].
\end{displaymath}

Поэтому вместо исходной задачи минимизации функции $\varphi(y)$ в области $D$ можно рассмотреть минимизацию одномерной функции $f(x)=\varphi(y(x)) $, удовлетворяющий условию Гельдера для $ x\in [0,1]$.

Чтобы определить, в какой момент лучше запускать локальный метод в рамках глобального поиска, воспользуемся алгоритмом, основанным на дереве решений. На первом этапе поиска происходит накопление поисковой информации в виде результатов выполненных испытаний. На основе результатов испытаний проводилось обучение дерева решений, позволяющего получать кусочно-постоянную аппроксимацию целевой функции и на ее основе прогнозировать поведение целевой функции.

Дерево решений — это инструмент для автоматизированного анализа больших массивов данных, применяемый в машинном обучении. Его можно использовать как для решения задач классификации, так и для регрессии. В случае построения регрессии каждому листу дерева присваивается константа. Поэтому полученная аппроксимирующая функция является кусочно-постоянной. В нашем случае дерево решений строит функцию $\psi(y)$ как кусочно-постоянную аппроксимацию $\varphi(y)$ в области поиска. Обозначим значение, вычисленное деревом решений в точке $y$, как $z' = \psi(y)$.
На рис. \ref{fig:fig2} представлен график целевой функции $\varphi(y)$; На рис. \ref{fig:fig2_2} представлена соответствующая аппроксимация $\psi(y)$, построенная с помощью дерева решений.

\begin{figure}
	\begin{center}
		\begin{minipage}[h]{0.7\linewidth}
			\includegraphics[width=1\linewidth]{figure/fig5.png}
			\caption{Целевая функция $\varphi(y)$ из тестового класса} %% подпись к рисунку
			\label{fig:fig2}
		\end{minipage}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\begin{minipage}[h]{0.7\linewidth}
			\includegraphics[width=1\linewidth]{figure/fig4.png}
			\caption{Кусочно-постоянная аппроксимация $\psi(y)$ функции $\varphi(y)$} %% подпись к рисунку
			\label{fig:fig2_2}
		\end{minipage}
	\end{center}
\end{figure}

При реализации алгоритма поиска областей притяжения локальных минимумов для построения дерева решений использовались алгоритмы из библиотеки OpenCV. OpenCV — это библиотека с открытым исходным кодом алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения. Более подробную информацию о дереве решений можно найти в \cite{Brahmbhatt2013}.

После определения с помощью дерева решений области притяжения локального минимума, из ближайшей точки запускаем локальный метод. Алгоритм поиска локального минимума Хука--Дживса относится к классу методов нулевого порядка. Его правила вычисления представляют собой комбинацию исследовательского поиска (для выбора направления) и поиска в выбранном направлении \cite{Himmelblau72}.

\begin{figure} 
	\begin{center} 
		\begin{minipage}[h]{0.8\linewidth} 
			\includegraphics[width=1\linewidth]{figure/fig1.pdf} 
			\caption{Пример итераций алгоритма Хука-Дживса} %%  подпись к рисунку 
			\label{fig:fig1} 
		\end{minipage} 
	\end{center} 
\end{figure}

На рис. \ref{fig:fig1} показан пример работы алгоритма. Отображаются линии  уровня, закрашенные кружки обозначают успешные шаги, незакрашенные кружки соответствуют неудачным шагам, выполненным во время исследовательского поиска.

\section{Многомерный параллельный алгоритм глобального поиска}\label{SecGSA}

На начальном этапе мастер-процесс (пусть он имеет идентификатор $0$) параллельно запускает $p$ испытаний в $p$ разных точках $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$ отрезка $[0,1]$.
Прообразами двух точек являются граничные $x^1=0,x^p=1$; остальные — внутренние точки $x^i\in\left(0,1\right),i=2,\ldots,p-1$.

Теперь предположим, что алгоритм уже сделал $K\geq0$ итераций и проведено $k$ испытаний.

\begin{enumerate}

	\item Извлекаем из множество $U$ (координат точек испытания) $c = max(p, |U|)$ координат точек испытания $y^{k+1}=y\left(x^{k+1}\right) ... y^{k+c}=y\left(x^{k+c}\right)$. Если $c < p$ то сначала вычисляем $n = p - c$ точек согласно шагам 2 - 7, иначе сразу переходим к шагу 8.

	\item Перенумеруем (по нижними индексами) точки $x^i, 1\leq i\leq k$, а также граничные точки отрезка $[0,1]$ в порядке возрастания координаты  
	\begin{equation} 
		\label{agp1_sort} 	0=x_0<\ x_1<\ ...\ <x_{k+1}=1. 	
	\end{equation} 
	и сопоставить их со значениями $z_i=f(x_i)$.

	\item  Вычислить текущую нижнюю оценку $M$ неизвестной константы Гельдера $H$: 
	\begin{equation} 
		\label{agp2_mu} 	\mu=max\left\{\frac{|z_i-z_{i-1}|}{{{(x}_i-x_{i-1})}^{1/N}},\ i=1,\ldots,k\right\},\ M=\  \left\{\begin{matrix}r\mu,\ \mu>0,\\1,\ \mu=0,\\\end{matrix}\right.\ 	
	\end{equation} 
	где $r>1$ — параметр алгоритма. Этот параметр управляет надежностью алгоритма: более высокие значения $r$ гарантируют гарантированное нахождение глобального минимума, выбор меньшего значения ускоряет сходимость алгоритма.

	\item  Для каждого интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить значение $R(i)$, называемое \textit{характеристикой} интервала, по формулам

	\begin{equation} 
		\label{agp3_R1} R(1)=2\Delta_1-4\dfrac{z_1}{M}, \; R(k+1)=2\Delta_{k+1}-4\dfrac{z_k}{M}, 
	\end{equation}

	\begin{equation} 
		\label{agp3_Ri} R(i)=\Delta_i+\dfrac{(z_i-z_{i-1})^2}{M^2\Delta_i}-2\dfrac{z_i+z_{i-1}}{M},1<i<k+1, 
	\end{equation}

	где \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).

	\item   Упорядочить характеристики $R\left(i\right),\ 1\leq i \leq k+1,$ в порядке не возрастания
	\begin{equation} 
		\label{agp4_R_sort} 	R\left(t_1\right)\geq\ R\left(t_2\right)\geq...\geq\ R\left(t_k\right)\geq\ R(t_{k+1}),\  
	\end{equation} 
	и выделить $n$ интервалов с индексами $t_j,\ 1\le\ j\le\ n$  с наибольшими значениями характеристик.

	\item Вычислить новые координаты точки испытания $y'^{k+j}=y\left(x^{k+j}\right), \ 1\leq j\leq n$ прообразом которого является $x^{k+j}\in\left(x_{t_j-1},x_{t_j}\right)$, в соответствии с формулой
	\begin{equation}
		\label{agp5_x1}
		x^{k+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}\left(z_t-z_{t-1}\right)\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{M}\right]^N.
	\end{equation}

	\item Отправить на процесс $j, \ 1\leq j\leq n$ координату $y'^{k+1} ... y'^{k+n}$  и сообщаем процессу, что нужно произвести одиночное испытание.

	\item Отправить на процессы $h, \ 1\leq h\leq c$ координаты точек $y^{k+1} ... y^{k+c}$ и сообщаем процессу, что нужно запустить локальный метод.

	\item Получить со всех $p$ процессов значения испытаний и добавить их в поисковую информацию.

	\item Добавить точки испытаний $y'^{k+1}=y\left(x^{k+1}\right) ... y'^{k+n}=y\left(x^{k+т}\right)$ к множеству $V$.

	\item Если $k\ <\ 100 N$, перейти к шагу 1.

	\item Строим дерево решений по всей накопленной поисковой информации, получаем аппроксимирующую функцию $\psi(y)$.

	\item Если дерево решений используется впервые, постройте равномерную сетку
	\begin{displaymath} 
		Y'=\{ y'\in R^N:\ a_i\le y'_i \le b_i,\ 1\le i\le N \},
	\end{displaymath} 
	где количество узлов сетки является параметром метода.

	\item Вычислить кусочно постоянную аппроксимацию функции $\psi(y)$, по дереву решений : $Z' = \{ z'=  \psi(y'), y' \in Y'\}$

	\item Для всех точек $y'\in V$ найдите точки $y'_q$, ближайшие к $y'$,

	Если ни одна точка $y'_q$ не имеет меньшего значения, чем $y'_q$, поместить $y'$ в множество $U$

	Очистить множество $V$.

	\item Проверка критериев остановки

\end{enumerate}

Следующие $p$ испытания выполняются параллельно в точках $x^{k+j},\ 1\leq j\leq p$, вычисляемых по формуле (\ref{agp5_x1}). По завершении испытаний результаты этих испытаний сохраняются в базе поисковой информации, а алгоритм переходит к вычислению новых точек испытания.
Отметим, что, как правило, процесс испытания в прикладных оптимизационных задачах гораздо более затратный по вычислительным ресурсам по сравнению с вычислением координаты точки.

Алгоритм останавливается в том случае, если выполняется условие \(\Delta_{t_j} < \varepsilon\) хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$, из (\ref{agp4_R_sort} ). Этот критерий остановки (наряду с обычным для итерационных методов критерием ограничения числа выполняемых итераций) используется в прикладных задачах оптимизации, в которых точка глобального минимума $y^*$ априори неизвестна.

При решении тестовых задач, в которых известна точка глобального минимума $y^*$, можно использовать критерий остановки и по попаданию в окрестность глобального минимума. В этом случае метод останавливается, если выполняется условие $\left\|y(x_{t_j})-\ y^\ast\right\| < \varepsilon$ для одного $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}).

В качестве окончательной оценки глобального минимума рассматриваемой задачи берется значение 
\begin{equation} 
	f_k^*=\min_{1\leq i \leq k}f(x_i), \; x_k^*=arg \min_{1\leq i \leq k}f(x_i). 
\end{equation}

Обоснование такого способа организации вычислений см. в \cite{Strongin2000,Barkalov2016}. Модификации с учетом наличия ограничений-неравенств, а также информация о производной целевой функции представлены в \cite{Barkalov2002,Gergel1997}.

% переводить до этого раздела
\section{Численные эксперименты}\label{SecR}

Численные эксперименты проводились на суперкомпьютере Лобачевского. Каждый узел имел по два процессора Intel Sandy Bridge E5-2660 2,2 ГГц, 64 Гб оперативной памяти.

В экспериментах использовался генератор тестовых задач GKLS, который может генерировать задачи многоэкстремальной оптимизации с известными свойствами: точка глобального минимума, количество локальных минимумов и др.

Ниже представлены результаты работы синхронного параллельного алгоритма глобального поиска  с использованием дерева решений для нахождения областей притяжения локальных минимумов. Численное сравнение проводилось на двух классах функций GKLS (Simple и Hard из \cite{Sergeyev2006}) размерностей 2, 3, 4 и 5.
Эти два класса отличаются размером области притяжения точки глобального минимума.
Для простого класса задач радиус области притяжения в три раза больше, чем для сложного.

Критерием остановки алгоритмов было попадание очередных пробных точек в $\varepsilon$-близость к истинному глобальному минимуму. В таблицах \ref{tab:1} \ref{tab:2} сравниваются усредненные числа итераций, выполняемых алгоритмами при решении задач, и значения ускорения времени относительно последовательного прогона. Распараллеливание выполнено по технологии MPI, эксперимент выполнен на 8, 16, 32 и 64 процессах.

\begin{table}[!ht]
	\caption{Ускорение, выполненных разными алгоритмами}
	\label{tab:2}
	\centering	
	\begin{tabular}{|l|l|l|l|l|}
		\hline
		2 & 8 mpi & 16 mpi & 32 mpi & 64 mpi  \\ \hline
		hard & 6,416962112 & 7,429653192 & 16,44977408 & 7,698208433  \\ \hline
		simple & 11,19015498 & 5,017982941 & 13,58978911 & 30,41682225  \\ \hline
		3 & ~ & ~ & ~ &   \\ \hline
		hard & 4,010618178 & 6,760486061 & 22,40788476 & 23,48393504  \\ \hline
		simple & 3,975141387 & 5,069355695 & 14,97196995 & 14,48568142  \\ \hline
		4 & ~ & ~ & ~ &   \\ \hline
		hard & 3,516700024 & 7,767619059 & 15,63797476 & 22,60877765  \\ \hline
		simple & 4,204314065 & 6,771626114 & 7,265190222 & 5,789337963  \\ \hline
		5 & ~ & ~ & ~ &   \\ \hline
		hard & 10,28406793 & 18,45856679 & 38,75628766 & 30,57030882  \\ \hline
		simple & 10,79552171 & 15,93214599 & 31,96240356 & 31,65420434  \\ \hline
	\end{tabular}
\end{table}

Как видно из таблицы \ref{tab:2}, ускорение итерации достаточно велико для задач любой размерности.

\section{Conclusion}

На данный момент в результате проделанной работы удалось совместить алгоритм глобального поиска с методом локальной оптимизации. В отличие от известных многозаходных схем решение о запуске локального метода принимается с помощью дерева решений. Использование такой комбинации методов позволяет значительно ускорить работу алгоритма.

Параллельная версия алгоритма сохраняет свойства своего последовательного прототипа, что было подтверждено численными экспериментами по решению серии из нескольких сотен задач различной размерности. Предлагаемая схема позволяет использовать преимущества как распараллеливания, так и быстрого поиска локальных экстремумов.

Помимо использования деревьев решений для выявления областей притяжения локальных экстремумов многоэкстремальных функций, мы также планируем использовать методы машинного обучения для разделения переменных решаемой задачи. Во многих прикладных задачах оптимизации зависимость целевой функции от некоторых параметров бывает либо линейной, либо одномодальной. Выделение таких переменных в специальную группу и решение задачи с помощью параллельной схемы рекурсивной оптимизации \cite{Barkalov2020_1} позволяет сократить время решения задачи на порядки по сравнению с использованием глобального поиска сразу по всем переменным.

%
% ---- Bibliography ----
%
\bibliographystyle{spmpsci}
\bibliography{bibliography}{}

\end{document}

### 2025-07-20 2025/Problems of Informatics/article.tex

@@
 \begin{small}
-% аннотация рус.
-В статье обсуждается проблема поиска глобального минимума функции, которая задана в виде <<черного ящика>> и может быть не всюду определена в области поиска.
-Такие функции возникают при решении прикладных задач из-за особенностей оптимизируемого объекта или метода моделирования.
-Например, численная нестабильность метода при определенном сочетании параметров может привести к возникновению подобластей, в которых невозможно корректно провести моделирование и оценить значение целевой функции.
-В некоторых случаях эти подобласти известны, но в большинстве случаев информация о них отсутствует.
-Существование подобластей, в которых целевая функция является неопределенной, можно интерпретировать как наличие некоторых скрытых, заранее неизвестных ограничений задачи.
-Предлагаемый подход к решению такого рода задач является расширением информационно-статистического алгоритма глобального поиска и учитывает возможность неопределенных значений целевой функции в некоторых точках.
-В рамках предложенного алгоритма на каждом шаге поиска проводится оценка характеристик поисковых интервалов на основе значений целевой функции, вычисленных на их границах. 
-В случае отсутствия информации о значениях функции для вычисления характеристики интервала используются оценка, учитывающая длину этого интервала.
-Для сокращения количества испытаний в подобластях, в которых функция не определена, предлагается использовать специальный параметр метода, позволяющий регулировать число точек испытаний в области невычислимости.
-В статье изложено описание модифицированного алгоритма глобального поиска для решения такого класса задач.
-Приведены результаты экспериментов на серии тестовых задач.
-Произведено сравнение с другими известными алгоритмами глобальной оптимизации.
-Приводится пример решения прикладной задачи настройки гиперпараметров метода машинного обучения, в которой настраиваемый метод работает некорректно при некоторых сочетаниях его параметров.
+% новая аннотация рус.
+В статье рассматриваются задачи поиска наилучшего сочетания гиперпараметров методов машинного обучения и искусственного интеллекта. 
+В таких задачах актуальной является проблема некорректной работы исследуемых методов ИИ и МО в некоторых (заранее неизвестных) подобластях области изменения гиперпараметров. 
+С математической точки зрения такая задача может быть представлена как задача поиска глобального минимума функции, заданной в виде <<черного ящика>> и не всюду определенной в области поиска. 
+Существование подобластей, где целевая функция является неопределенной, можно интерпретировать как наличие некоторых скрытых, заранее неизвестных ограничений.
+Предложен подход к решению такого рода задач, который является расширением информационно-статистического алгоритма глобального поиска и учитывает наличие неопределенных значений целевой функции в некоторых точках.
+В рамках предложенного алгоритма проводится разбиение области поиска точками испытаний и оцениваются характеристики подобластей на основе значений целевой функции, вычисленных на их границах. 
+В случае отсутствия информации о значениях функции в алгоритме используются оценка, учитывающая размер исследуемой подобласти. 
+Для сокращения количества испытаний в подобластях, в которых функция не определена, введен специальный параметр метода, позволяющий регулировать число точек испытаний в области невычислимости.
+Изложено подробное описание и приведена схема работы модифицированного алгоритма глобального поиска.
+Продемонстрированы результаты его сравнения с другими известными алгоритмами глобальной оптимизации, полученные при проведении численных экспериментов как с тестовыми функциями, так и модельными задачами настройки гиперпараметров, в которых возникают неопределенные значения оптимизируемой метрики качества.
 
@@
 При этом программные реализации настраиваемых методов могут вести себя абсолютно по-разному: в лучшем случае информируют пользователя о том, что задача не может быть решена, в худшем -- возвращают недопустимое значение (NaN, inf) в качестве найденного решения.
-Проблема возникновения недопустимых комбинаций, которая не имеет особого значения в случае <<ручной>> настройки, стала камнем преткновения для многих фреймворков автоматической настройки гиперпараметров.
+Проблема возникновения недопустимых комбинаций, которая не имеет особого значения в случае <<ручной>> настройки, существенно осложняет работу фреймворков автоматической настройки гиперпараметров.
 
@@
 Настоящая работа продолжает развитие одного из эффективных детерминированных методов решения задач липшицевой глобальной оптимизации -- ин\-фор\-ма\-ци\-он\-но-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска [\ref{rfa:rulit:Sergeyev2013},\ref{rfa:rulit:Strongin2000}].
-В статье приведено описание алгоритма глобального поиска, адаптированного для работы с частично определенной целевой функцией.
+В статье приведено описание нового алгоритма, адаптированного для работы с частично определенной целевой функцией.
 Данный алгоритм основан на подходе, предложенном авторами ранее в [\ref{rfa:rulit:Usova2024}].
 Проведено исследование поведения других известных алгоритмов при решении задач с не всюду вычислимой целевой функцией.
-Продемонстрированы результаты численных экспериментов как с тестовыми задачами, так и несколькими прикладными задачами настройки гиперпараметров, в которых возникают недопустимые сочетания гиперпараметров.
-Выполнена настройка метода LinearSVC при решении задачи классификации на известном наборе данных Iris, а также метода предсказания значений временного ряда фреймворка FEDOT для датасета montly beer prodaction.
+Продемонстрированы результаты численных экспериментов как с тестовыми функциями, так и модельными задачами настройки гиперпараметров, в которых возникают неопределенные значения целевой функции.
+А именно, проведена настройка гиперпараметров метода LinearSVC при решении задачи классификации на наборе данных Iris, а также метода предсказания значений временного ряда для датасета montly beer production с платформы Kaggle.
 
@@
 \begin{algorithm}
-\caption{Алгоритм глобального поиска на случай не всюду вычислимой целевой функции (АГП-Н)}\label{main_alg}
+\caption{Алгоритм глобального поиска для не всюду вычислимой целевой функции (АГП-Н)}\label{main_alg}
 \begin{algorithmic}[1]
@@
 В качестве второй модельной задачи рассмотрим задачу настройки гиперпараметров методов предсказания значений временного ряда (time series forecasting).
-Предсказание значений производилось с использованием фреймворка FEDOT %\cite{FEDOT,nikitin2022automated}
+Предсказание значений производилось с использованием фреймворка FEDOT [\ref{nikitin2022automated}]
 (фреймворк с открытым исходным кодом для задач автоматизированного моделирования и машинного обучения, AutoML).
@@
 
-Исследование проводилось на датасете \textit{montly beer prodaction}, который содержит 476 записей валидационной выборки. %\cite{MonthlyBeerDataset,MonthlyBeerArticle}
-
+Исследование проводилось на датасете \textit{montly beer production}, который содержит 476 записей. %\cite{MonthlyBeerDataset,MonthlyBeerArticle}
+Указанный датасет представлен на платформе Kaggle и часто используется для тестирования алгоритмов машинного обучения (см., например, [\ref{MonthlyBeerArticle}]).
 Построенный для решения поставленной задачи в FEDOT пайплайн включает в себя комбинацию из двух преобразований: lagged и cgru.
@@
 \centering
-\caption{Результаты решения задачи настройки гиперпараметров метода предсказания значений временного ряда на датасете montly beer prodaction с использованием фреймворков iOpt и Optuna}\label{tab1}
+\caption{Результаты решения задачи настройки гиперпараметров метода предсказания значений временного ряда на датасете montly beer production с использованием фреймворков iOpt и Optuna}\label{tab1}
 \begin{tabular}{|l|c|c|}
@@
 
-В настоящей работе были рассмотрены актуальные проблемы решения задач настройки гиперпараметров методов искусственного интеллекта (ИИ) и машинного обучения (МО).
-
-Основная сложность таких задач заключается в довольно длительном решении соответствующей исходной задачи для каждого выбранного сочетания гиперпараметров, что накладывает ограничение на возможное количество испытываемых комбинаций.
-
-В свою очередь принципиально новой и не менее сложной проблемой является некорректная работа исследуемых алгоритмов ИИ и МО в некоторых заранее неизвестных подобластях области изменения гиперпараметров; фактически, область допустимых сочетаний параметров является заранее неопределенной.
-Проблема возникновения таких недопустимых комбинаций, которая ранее не имела особого значения при <<ручной>> настройке, стала камнем преткновения для многих фреймворков автоматической настройки гиперпараметров.
-
-Для решения исследуемого класса задач была разработана модификация одного из эффективных детерминированных методов решения задач липшицевой глобальной оптимизации -- ин\-фор\-ма\-ци\-он\-но-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска.
-Разработанный алгоритм реализован на базе фреймворка методов интеллектуальной оптимизации с открытым исходным кодом iOpt.
-
-В статье изложено подробное описание модифицированного алгоритма (АГП-Н).
-Приведены результаты серийных экспериментов, демонстрирующие надёжность разработанного подхода.
-Продемонстрировано сравнение результатов работы АГП-Н с результатами, полученными с помощью стандартных алгоритмов оптимизации из библиотеки SciPy.
-Показана эффективность разработанного метода при решении исследуемого класса задач.
-
-В качестве демонстрационных примеров были решены несколько задач настройки гиперпараметров, в которых возникает проблема недопустимых сочетаний гиперпараметров, приводящих к <<падению>> метода.
-
-В качестве первой модельной задачи была рассмотрена задача настройки гиперпараметров классического алгоритма классификации LinearSVC.
-Разработанный метод успешно решил поставленную задачу, найдя лучшее сочетание гиперпараметров по сравнению со стандартными алгоритмами из библиотеки scikit-learn.
-
-В качестве второй модельной задачи была рассмотрена задача предсказания значений временного ряда средствами фреймворка FEDOT.
-В ходе эксперимента также было проведено сравнение с другим известным фреймворком, применяемым для настройки гиперпараметров -- Optuna.
-В ходе эксперимента фреймворк iOpt продемонстрировал свое преимущество во времени решения задачи, немного уступив в качестве.
+В настоящей работе рассматривалась актуальная проблема решения задач настройки гиперпараметров методов искусственного интеллекта (ИИ) и машинного обучения (МО).
+Сложность таких задач заключается в большой трудоемкости оценки метрик качества методов ИИ и МО для заданного сочетания гиперпараметров, что ограничивает возможное число проверяемых комбинаций.
+Одновременно с этим исследуемые методы ИИ и МО могут работать некорректно в некоторых заранее неизвестных подобластях всей области изменения гиперпараметров; фактически, область допустимых сочетаний гиперпараметров является неопределенной.
+Проблема возникновения некорректных сочетаний гиперпараметров, которая ранее не имела особого значения при <<ручной>> настройке, при использовании автоматической настройки стала весьма актуальной.
+
+Для решения указанного класса задач была разработана модификация одного из эффективных детерминированных методов решения задач глобальной оптимизации -- ин\-фор\-ма\-ци\-он\-но-ста\-тис\-ти\-чес\-ко\-го алгоритма глобального поиска.
+Предложенный алгоритм глобального поиска для не всюду вычислимых функций (АГП-Н) реализован на базе фреймворка методов интеллектуальной оптимизации с открытым исходным кодом iOpt.
+
+В статье подробно описаны вычислительные правила и приведена схема работы АГП-Н.
+Представлены результаты экспериментов на нескольких сотнях тестовых задач, демонстрирующие надежность предложенного подхода к обработке неопределенных значений оптимизируемой функции.
+Также приведены результаты сравнения работы АГП-Н и стандартных алгоритмов оптимизации из библиотеки SciPy, которые подтверждают эффективность разработанного метода при решении тестового класса задач.
+
+В качестве модельных примеров были решены несколько задач настройки гиперпараметров, где возникает проблема возникновения недопустимых сочетаний гиперпараметров, при которых невозможно оценить целевую метрику качества.
+В первом примере была проведена настройка гиперпараметров классического алгоритма классификации LinearSVC.
+АГП-Н успешно решил поставленную задачу, найдя лучшее сочетание гиперпараметров по сравнению со стандартными алгоритмами из библиотеки scikit-learn.
+В качестве второй модельной задачи была рассмотрена настройка гиперпараметров метода предсказания значений временного ряда.
+В ходе эксперимента было проведено сравнение АГП-Н с алгоритмами настройки из широко известного фреймворка Optuna.
+АГП-Н показал сравнимый результат (в терминах полученного значения целевой метрики), значительно обогнав Optuna по времени, затраченному на решение задачи.
 

### 2025-07-25 2025/Problems of Informatics/article.tex
new eng annotation added and references fixed

@@
 % аннотация англ.
-\colorbox{orange}{исправить в соответсвии с введением и экспериментами ->}
-The paper discusses the problem of finding a global minimum of a function that may be multiextremal, non-differentiable, given in the form of a ``black box'' and computable only in some part of the search domain.
+The article discusses  the problem of finding the best combination of hyperparameters of machine learning and artificial intelligence methods.
+In many cases, the efficiency (in a given metric) of the resulting solution for different values of hyperparameters can be quite different.
+В таких задачах актуальной является проблема некорректной работы исследуемых artificial intelligence (AI) and machine learning (ML) methods в некоторых (заранее неизвестных) подобластях области изменения гиперпараметров.
+При этом вычислительная сложность настройки делает невозможным использование ручного или полного перебора.
+Данные особенности определили необходимость разработки различных методов интеллектуального автоматического подбора гиперпараметров.
+С математической точки зрения такая задача может быть представлена как problem of finding a global minimum of a function, given in the form of a ``black box'' and computable only in some part of the search domain.
 In this case, each computation of the objective function value at some point of the feasible domain may require significant computing resources.
@@
 The existence of subdomains where the objective function is undefined can be interpreted as the existence of some hidden, a priori unknown constraints of the problem.
-Such an effect can occur when solving applied problems due to the specific features of the optimized object or modeling method (for example, numerical instability of the method for some combination of parameters).
-In such subdomains, it is impossible to correctly perform numerical modeling and estimate the value of the objective function.
-A special case is the problem of tuning hyperparameters of artificial intelligence (AI) and machine learning (ML) methods.
-In these problems, the efficiency (in a given metric) of the resulting solution for different values of hyperparameters can be quite different.
-Moreover, the computational complexity of tuning makes the exhaustive search practically inapplicable.
-These features required the development of various methods for intelligent automatic tuning of hyperparameters.
-At the same time, the problem of existence of unacceptable combinations of hyperparameters, which was not particularly important in the case of manual tuning, has become a stumbling block for many frameworks using intelligent tuning methods.
-Our goal is to develop an algorithm in which the absence of a value of the objective function at some point should not lead to its incorrect operation.
-The proposed approach to solving such problems extends the information-statistical global search algorithm (GSA).
-This algorithm solves multidimensional problems by reducing them to one-dimensional optimization problems using space-filling curves (Peano curves).
-In this study, new rules for processing search intervals with no objective function values at their boundary points were proposed.
-The calculation of the characteristic of an interval with two non-computable points is based on the interval length.
-To minimize the number of redundant trials in subdomains where the function is not defined, we used a special method parameter $\alpha$ that regulates the number of trial points in these subdomains.
-The calculation of the characteristic of an interval with non-computable and computable points is performed according to the rules for working with a boundary interval.
-The article provides a detailed description of a modified global search algorithm for solving the class of problems under consideration.
+The authors propose an approach to solving this type of problem, , which is an extension of the information-statistical global search algorithm (GSA) and takes into account the presence of undefined values of the objective function at some points.
+В рамках алгоритма происходит разбиение области поиска точками испытаний и оцениваются характеристики подобластей на основе значений целевой функции, вычисленных на их границах.
+В случае отсутствия информации о значении функции в точке в алгоритме используется предложенная авторами оценка, учитывающая размер исследуемой подобласти.
+To minimize the number of redundant trials in subdomains where the function is not defined, we used a special method parameter $\alpha$ that regulates the number of trial points in areas of non-computability.
+The solution of multidimensional problems implemented through reducing them to one-dimensional optimization problems using space-filling curves (Peano curves).
+The article provides a detailed description and a diagram of the operation of the modified global search algorithm.
 The implementation of the global search algorithm for the case of a not-everywhere computable objective function (GSA-N) was based on the iOpt open source framework of intelligent optimization methods.
@@
 By adjusting the parameter $\alpha$, it was possible to achieve the same speed of GSA-N operation as the basic GSA.
-The paper also considered the behavior of global optimization algorithms of the \textit{scipy.optimize} library when solving problems with non-computable domains.
+The paper also considered the behavior of global optimization algorithms of the scipy.optimize library when solving problems with non-computable domains.
 In this study, the differential evolution and brute force methods showed the worst results, failing to solve this type of problem at all.
 The DIRECT and SHGO methods, although they solved the problems, were less effective than the developed GSA-N.
-In addition, the paper presents the results of numerical experiments obtained when tuning the hyperparameters of the LinearSVC method used to solve the classification problem on the well-known Iris dataset.
+Основным результатом являются эксперименты с модельными задачами настройки гиперпараметров, в которых возникают неопределенные значения оптимизируемой метрики качества.
+В этих задачах существуют недопустимые сочетания гиперпараметров, при которых настраиваемый метод возвращает бесконечное значение целевой функции.
+Так была успешно проведена настройка классического алгоритма классификации LinearSVC.
+АГП-Н успешно решил поставленную задачу, найдя лучшее сочетание гиперпараметров по сравнению со стандартными алгоритмами из библиотеки scikit-learn.
+Также был настроен метода предсказания значений временного ряда на базе фреймворка FEDOT.
+В ходе этого эксперимента также было проведено сравнение АГП-Н с алгоритмами настройки из другого широко известного фреймворка Optuna.
+АГП-Н показал сравнимый результат (в терминах полученного значения целевой метрики), значительно обогнав Optuna по времени, затраченному на решение задачи.
 
 %Here is the text of the extended annotation in English, ranging from 4000 to 8000 characters.
-%В статье обсуждается проблема поиска глобального минимума функции, которая может быть многоэкстремальной, недифференцируемой, более того, заданной в форме «черного ящика» и вычислимой лишь в некоторой части области поиска.
+
+%В статье рассматриваются задачи поиска наилучшего сочетания гиперпараметров методов машинного обучения и искусственного интеллекта.
+Во многих случаях разница в эффективности (в заданной метрике) получаемого решения при разных значениях гиперпараметров может быть достаточно значительной.
+%В таких задачах актуальной является проблема некорректной работы исследуемых методов ИИ и МО в некоторых (заранее неизвестных) подобластях области изменения гиперпараметров.
+При этом вычислительная сложность настройки делает невозможным использование ручного или полного перебора.
+Данные особенности определили необходимость разработки различных методов интеллектуального автоматического подбора гиперпараметров.
+%С математической точки зрения такая задача может быть представлена как задача поиска глобального минимума функции, заданной в виде <<черного ящика>> и не всюду определенной в области поиска.
 При этом каждое вычисление функции в некоторой точке допустимой области может требовать значительных вычислительных ресурсов.
 О целевой функции известно лишь что она удовлетворяет условию Липшица.
-%Наличие подобластей, в которых целевая функция является неопределенной, можно интерпретировать как наличие некоторых скрытых, заранее неизвестных ограничений задачи.
-Такой эффект может возникать при решении прикладных задач в силу особенностей оптимизируемого объекта или метода моделирования (например, численная нестабильность метода при определенном сочетании параметров).
-В таких подобластях невозможно корректно провести численное моделирование и оценить значение целевой функции.
-Частным случаем является задача настройки гиперпараметров методов искусственного интеллекта (ИИ) и машинного обучения (МО).
-Во многих случаях разница в эффективности (в заданной метрике) получаемого решения при разных значениях гиперпараметров может быть достаточно значительной.
-При этом вычислительная сложность настройки делает невозможным использование полного перебора.
-Данные особенности определили необходимость разработки различных методов интеллектуального автоматического подбора гиперпараметров.
-При этом проблема возникновения недопустимых комбинаций гиперпараметров, которая не имела особого значения в случае «ручной» настройки, для многих фреймворков, использующих методы интеллектуальной настройки, стала камнем преткновения.
-%Нашей целью является разработка алгоритма, в котором отсутствие значения целевой функции в некоторой точке не должно приводить к его некорректной работе.
-Предлагаемый подход к решению такого рода задач является расширением информационно-статистического алгоритма глобального поиска (АГП).
-Данный алгоритм предполагает решение многомерных задач посредством их сведения к задачам одномерной оптимизации при помощи кривых, заполняющих пространство (кривых Пеано).
-В рамках данного исследования были предложены новые правила обработки поисковых интервалов, в граничных точках которых отсутствуют значения целевой функции.
-Расчёт характеристики интервала с двумя невычислимыми точками производится, основываясь на длине интервала.
-Для минимизации количества избыточных испытаний в областях, в которых функция не определена, при таком подходе используется специальный параметр метода, позволяющий регулировать число точек испытаний в области невычислимости.
-Расчёт характеристики интервала с невычислимой и вычислимой точками производится по правилам работы с граничным интервалом.
-В статье изложено подробное описание модифицированного алгоритма глобального поиска для решения исследуемого класса задач.
+Существование подобластей, где целевая функция является неопределенной, можно интерпретировать как наличие некоторых скрытых, заранее неизвестных ограничений.
+%Авторами предложен подход к решению такого рода задач, который является расширением информационно-статистического алгоритма глобального поиска (АГП) и учитывает наличие неопределенных значений целевой функции в некоторых точках.
+В рамках алгоритма происходит разбиение области поиска точками испытаний и оцениваются характеристики подобластей на основе значений целевой функции, вычисленных на их границах.
+В случае отсутствия информации о значении функции в точке в алгоритме используется предложенная авторами оценка, учитывающая размер исследуемой подобласти.
+Для минимизации количества избыточных испытаний в областях, в которых функция не определена, при таком подходе используется специальный параметр метода $\alpha$, позволяющий регулировать число точек испытаний в области невычислимости.
+%Решение многомерных задач осуществляется посредством их сведения к задачам одномерной оптимизации при помощи кривых, заполняющих пространство (кривых Пеано).
+%В статье изложено подробное описание и приведена схема работы модифицированного алгоритма глобального поиска. 
 %Программная реализация алгоритма глобального поиска на случай не всюду вычислимой целевой функции (АГП-Н) была выполнена на базе фреймворка методов интеллектуальной эвристической оптимизации iOpt открытым исходным кодом.
-%Для проведения экспериментов был разработан генератор задач со скрытыми ограничениями GKLSHiddenConstraint, взявший за основу генератор GKLS, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами.
+%Для проведения численных экспериментов был разработан генератор тестовых задач со скрытыми ограничениями GKLSHiddenConstraint, взявший за основу генератор GKLS, позволяющий порождать задачи многоэкстремальной оптимизации с заранее известными свойствами.
 Данные функции были дополнены сгенерированными областями невычислимости в форме эллипсоидов (координата центры и радиусы генерировались случайно).
-%Приведенные в данной статье результаты экспериментов на серии тестовых задач GKLSHiddenConstraint демонстрируют надёжность глобального поиска и эффективность разработанного алгоритма.
-За счёт регулировки параметра $\alpha$ получилось добиться такой же скорости сходимости АГП-Н, как у базового АГП.
-%Также в статье было изучено поведение алгоритмов глобальной оптимизации библиотеки \textit{scipy.optimize} при решении задач с невычислимыми областями.
+%Полученные при проведении численных экспериментов с тестовыми функциями результаты демонстрируют надёжность и эффективность реализованного алгоритма.
+За счёт регулировки параметра $\alpha$ получилось добиться такой же скорости сходимости АГП-Н, как у АГП.
+%Также в статье было исследовано поведение известных алгоритмов глобальной оптимизации библиотеки \textit{scipy.optimize} при решении задач с невычислимыми областями.
 В рамках данного исследования методы differential\_evolution и brute показали наихудшие результаты, не справившись с решением такого вида задач совсем.
 Методы direct и shgo хоть и продемонстрировали хорошие результаты, оказались менее эффективны чем разработанный АГП-Н.
-%Помимо этого в статье продемонстрированы результаты численных экспериментов с прикладной задачей настройки гиперпараметров метода LinearSVC при решении задачи классификации на известном наборе данных Iris.
-<РЕЗУЛЬТАТЫ>.
-
-\textbf{Keywords:} Lipschitz global optimization, multi-extremal functions, black-box functions, partially defined functions, hidden constraints, hyperparameter tuning.
-\colorbox{orange}{<-}
+%Основным результатом являются эксперименты с модельными задачами настройки гиперпараметров, в которых возникают неопределенные значения оптимизируемой метрики качества.
+В этих задачах существуют недопустимые сочетания гиперпараметров, при которых настраиваемый метод возвращает бесконечное значение целевой функции.
+%Так была успешно проведена настройка классического алгоритма классификации LinearSVC.
+АГП-Н успешно решил поставленную задачу, найдя лучшее сочетание гиперпараметров по сравнению со стандартными алгоритмами из библиотеки scikit-learn.
+%Также был настроен метода предсказания значений временного ряда на базе фреймворка FEDOT.
+В ходе этого эксперимента также было проведено сравнение АГП-Н с алгоритмами настройки из другого широко известного фреймворка Optuna.
+АГП-Н показал сравнимый результат (в терминах полученного значения целевой метрики), значительно обогнав Optuna по времени, затраченному на решение задачи.
+
+\textbf{Keywords:} machine learning, hyperparameter tuning, global optimization, black-box functions, partially defined functions.
 \newline
@@
 
-%16
-\item \label{rfa:enlit:Sergeyev2013}
-Sergeyev~Y.D., Strongin~R.G., Lera~D. Introduction to Global Optimization Exploiting Space-Filling Curves. New York: Springer Briefs in Optimization, 2013. %DOI: 10.1007/978-1-4614-8042-6.
-
-%20
-\item \label{rfa:enlit:Strongin2000}
-Strongin~R.G., Sergeyev~Y.D. Global optimization with non-convex constraints. Sequential and parallel algorithms. Dordrecht: Kluwer Academic Publishers, 2000.
-
-%10
-\item \label{rfa:enlit:Kvasov2013}
-Kvasov~D.E., Sergeev~Ya.D. Metody lipshicevoj globalnoj optimizacii v zadachax upravleniya~// Avtomatika i telemexanika. 2013. No.~9. S.~3--19.
-
-%17
-\item \label{rfa:enlit:Sergeyev2020}
-Sergeyev~Y.D., Candelieri~A., Kvasov~D.E., Perego~R. Safe global optimization of expensive noisy black-box functions in the $\delta$-Lipschitz framework~// 
-Soft Comput. 2020. Vol.~24, No.~23. P.~17715--17735. %DOI: 10.1007/s00500-020-05030-3.
-
-%4
-\item \label{rfa:enlit:Dongarra2022}
-Dongarra~J.J. The evolution of mathematical software~// Commun. ACM. 2022. Vol.~65, No.~12. P.~66--72. %DOI: 10.1145/3554977.
-
-%5
-\item \label{rfa:enlit:Duwe2020}
-Duwe~K., et al. State of the art and future trends in data reduction for high-performance computing~// Supercomput. Front. Innov. 2020. Vol.~7, No.~1. %DOI: 10.14529/jsfi200101.
-
 %19
@@
 
+%16
+\item \label{rfa:enlit:Sergeyev2013}
+Sergeyev~Y.D., Strongin~R.G., Lera~D. Introduction to Global Optimization Exploiting Space-Filling Curves. New York: Springer Briefs in Optimization, 2013. %DOI: 10.1007/978-1-4614-8042-6.
+
+%20
+\item \label{rfa:enlit:Strongin2000}
+Strongin~R.G., Sergeyev~Y.D. Global optimization with non-convex constraints. Sequential and parallel algorithms. Dordrecht: Kluwer Academic Publishers, 2000.
+
+%%10
+%\item \label{rfa:enlit:Kvasov2013}
+%Kvasov~D.E., Sergeev~Ya.D. Metody lipshicevoj globalnoj optimizacii v zadachax upravleniya~// Avtomatika i telemexanika. 2013. No.~9. S.~3--19.
 \item \label{rfa:enlit:Usova2024}
@@
 
-%новые
-\item \label{rfa:enlit:differential_evolution}
-Storn~R., Price~K., Differential Evolution - a Simple and Efficient Heuristic for Global Optimization over Continuous Spaces // Journal of Global Optimization. 1997. Vol.~11. P.~341-359.
-
-\item \label{rfa:enlit:dual_annealing}
-Xiang~Y, Gubian~S, Suomela~B, Hoeng~J. Generalized Simulated Annealing for Efficient Global Optimization: the GenSA Package for R // The R Journal. 2013. Vol.~5, No.~1.
-
-\item \label{rfa:enlit:direct}
-Gablonsky~J., Kelley~C. A Locally-Biased form of the DIRECT Algorithm // Journal of Global Optimization. 2001. Vol.~21. P.~27--37.
-
-\item \label{rfa:enlit:basinhopping}
-Wales~D.J., Doye~J.P.K.
-Global Optimization by Basin-Hopping and the Lowest Energy Structures of Lennard-Jones Clusters Containing up to 110 Atoms // Journal of Physical Chemistry A.
-1997.
-Vol.~101.
-p.~5111.
-
-\item \label{rfa:enlit:shgo}
-Endres~S.C., Sandrock~C., Focke~W.W. A simplicial homology algorithm for lipschitz optimisation // Journal of Global Optimization. 2018.
-
 \item \label{rfa:enlit:Barkalov2021}
@@
 
+\item \label{rfa:enlit:iOptGithub}
+Isxodnyj kod frejmvorka iOpt -- URL: https://github.com/aimclub/iOpt (data obrashheniya: 26.01.2025).
+
 \item \label{rfa:enlit:iOptDocs}
 Dokumentaciya iOpt -- URL: https://iopt.readthedocs.io/ru/latest/ (data obrashheniya: 26.01.2025).
-
-\item \label{rfa:enlit:iOptGithub}
-Isxodnyj kod frejmvorka iOpt -- URL: https://github.com/aimclub/iOpt (data obrashheniya: 26.01.2025).
 
@@
 P.~469--480.
+
+\item \label{rfa:enlit:differential_evolution}
+Storn~R., Price~K., Differential Evolution - a Simple and Efficient Heuristic for Global Optimization over Continuous Spaces // Journal of Global Optimization. 1997. Vol.~11. P.~341-359.
+
+\item \label{rfa:enlit:dual_annealing}
+Xiang~Y, Gubian~S, Suomela~B, Hoeng~J. Generalized Simulated Annealing for Efficient Global Optimization: the GenSA Package for R // The R Journal. 2013. Vol.~5, No.~1.
+
+\item \label{rfa:enlit:direct}
+Gablonsky~J., Kelley~C. A Locally-Biased form of the DIRECT Algorithm // Journal of Global Optimization. 2001. Vol.~21. P.~27--37.
+
+\item \label{rfa:enlit:basinhopping}
+Wales~D.J., Doye~J.P.K.
+Global Optimization by Basin-Hopping and the Lowest Energy Structures of Lennard-Jones Clusters Containing up to 110 Atoms // Journal of Physical Chemistry A.
+1997.
+Vol.~101.
+p.~5111.
+
+\item \label{rfa:enlit:shgo}
+Endres~S.C., Sandrock~C., Focke~W.W. A simplicial homology algorithm for lipschitz optimisation // Journal of Global Optimization. 2018.
+
+\item \label{rfa:enlit:nikitin2022automated}
+Filippou~K., Aifantis~G., Papakostas~G.A., Tsekouras~G.E.
+Structure learning and hyperparameter optimization using an automated machine learning ({AutoML}) pipeline // Information.
+2023.
+Vol.~14, No.
+4.
+p.~232.
+
+\item \label{rfa:enlit:MonthlyBeerArticle}
+Xu~N. Time Series Analysis on Monthly Beer Production in Australia~// Highlights in Science, Engineering and Technology. 2024. Vol.~94. P.~392--401. DOI: 10.54097/4z3krj13
+
+%17
+%\item \label{rfa:enlit:Sergeyev2020}
+%Sergeyev~Y.D., Candelieri~A., Kvasov~D.E., Perego~R. Safe global optimization of expensive noisy black-box functions in the $\delta$-Lipschitz framework~// 
+%Soft Comput. 2020. Vol.~24, No.~23. P.~17715--17735. %DOI: 10.1007/s00500-020-05030-3.
+
+%4
+%\item \label{rfa:enlit:Dongarra2022}
+%Dongarra~J.J. The evolution of mathematical software~// Commun. ACM. 2022. Vol.~65, No.~12. P.~66--72. %DOI: 10.1145/3554977.
+
+%5
+%\item \label{rfa:enlit:Duwe2020}
+%Duwe~K., et al. State of the art and future trends in data reduction for high-performance computing~// Supercomput. Front. Innov. 2020. Vol.~7, No.~1. %DOI: 10.14529/jsfi200101.
+
+%\item \label{rfa:enlit:FEDOT}
+%Automated modeling and machine learning framework {FEDOT} -- URL:  https://github.com/aimclub/FEDOT (data obrashheniya: 25.07.2025).
 
@@
 В качестве второй модельной задачи рассмотрим задачу настройки гиперпараметров методов предсказания значений временного ряда (time series forecasting).
-Предсказание значений производилось с использованием фреймворка FEDOT [\ref{nikitin2022automated}]
+Предсказание значений производилось с использованием фреймворка FEDOT [\ref{rfa:rulit:nikitin2022automated}]
 (фреймворк с открытым исходным кодом для задач автоматизированного моделирования и машинного обучения, AutoML).
@@
 Исследование проводилось на датасете \textit{montly beer production}, который содержит 476 записей. %\cite{MonthlyBeerDataset,MonthlyBeerArticle}
-Указанный датасет представлен на платформе Kaggle и часто используется для тестирования алгоритмов машинного обучения (см., например, [\ref{MonthlyBeerArticle}]).
+Указанный датасет представлен на платформе Kaggle и часто используется для тестирования алгоритмов машинного обучения (см., например, [\ref{rfa:rulit:MonthlyBeerArticle}]).
 Построенный для решения поставленной задачи в FEDOT пайплайн включает в себя комбинацию из двух преобразований: lagged и cgru.
@@
 
-%16
-\item \label{rfa:rulit:Sergeyev2013}
-Sergeyev~Y.D., Strongin~R.G., Lera~D. Introduction to Global Optimization Exploiting Space-Filling Curves. New York: Springer Briefs in Optimization, 2013. DOI: 10.1007/978-1-4614-8042-6.
-
-%20
-\item \label{rfa:rulit:Strongin2000}
-Strongin~R.G., Sergeyev~Y.D. Global optimization with non-convex constraints. Sequential and parallel algorithms. Dordrecht: Kluwer Academic Publishers, 2000.
-
-%10
-\item \label{rfa:rulit:Kvasov2013}
-Квасов~Д.Е., Сергеев~Я.Д. Методы липшицевой глобальной оптимизации в задачах управления~// Автоматика и телемеханика. 2013. №~9. C.~3--19.
-
-%17
-\item \label{rfa:rulit:Sergeyev2020}
-Sergeyev~Y.D., Candelieri~A., Kvasov~D.E., Perego~R.
-Safe global optimization of expensive noisy black-box functions in the $\delta$-Lipschitz framework~// Soft Comput.
-2020.
-Vol.~24, No.~23.
-P.~17715--17735.
-DOI: 10.1007/s00500-020-05030-3.
-
-%4
-\item \label{rfa:rulit:Dongarra2022}
-Dongarra~J.J. The evolution of mathematical software~// Commun. ACM. 2022. Vol.~65, No.~12. P.~66--72. DOI: 10.1145/3554977.
-
-%5
-\item \label{rfa:rulit:Duwe2020}
-Duwe~K., et al. State of the art and future trends in data reduction for high-performance computing~// Supercomput. Front. Innov. 2020. Vol.~7, No.~1. DOI: 10.14529/jsfi200101.
-
 %19
@@
 Soft Comput. 2020. Vol.~24, No.~16. P.~11853--11865. DOI: 10.1007/s00500-020-05078-1.
+
+%16
+\item \label{rfa:rulit:Sergeyev2013}
+Sergeyev~Y.D., Strongin~R.G., Lera~D. Introduction to Global Optimization Exploiting Space-Filling Curves. New York: Springer Briefs in Optimization, 2013. DOI: 10.1007/978-1-4614-8042-6.
+
+%20
+\item \label{rfa:rulit:Strongin2000}
+Strongin~R.G., Sergeyev~Y.D. Global optimization with non-convex constraints. Sequential and parallel algorithms. Dordrecht: Kluwer Academic Publishers, 2000.
 
@@
 Commun. Comput. Inf. Sci. 2022. Vol.~1618. P.~293--306. DOI: 10.1007/978-3-031-11623-0{\_}20.
-
-%новые
-\item \label{rfa:rulit:differential_evolution}
-Storn~R., Price~K., Differential Evolution - a Simple and Efficient Heuristic for Global Optimization over Continuous Spaces //
-Journal of Global Optimization. 1997. Vol.~11. P.~341--359.
-
-\item \label{rfa:rulit:dual_annealing}
-Xiang~Y, Gubian~S, Suomela~B, Hoeng~J. Generalized Simulated Annealing for Efficient Global Optimization: the GenSA Package for R // The R Journal. 2013. Vol.~5, No.~1.
-
-\item \label{rfa:rulit:direct}
-Gablonsky~J., Kelley~C. A Locally-Biased form of the DIRECT Algorithm // Journal of Global Optimization. 2001. Vol.~21. P.~27--37.
-
-\item \label{rfa:rulit:basinhopping}
-Wales~D.J., Doye~J.P.K.
-Global Optimization by Basin-Hopping and the Lowest Energy Structures of Lennard-Jones Clusters Containing up to 110 Atoms // Journal of Physical Chemistry A.
-1997.
-Vol.~101.
-p.~5111.
-
-\item \label{rfa:rulit:shgo}
-Endres~S.C., Sandrock~C., Focke~W.W. A simplicial homology algorithm for lipschitz optimisation // Journal of Global Optimization. 2018.
 
@@
 P.~469--480.
+
+\item \label{rfa:rulit:differential_evolution}
+Storn~R., Price~K., Differential Evolution - a Simple and Efficient Heuristic for Global Optimization over Continuous Spaces //
+Journal of Global Optimization. 1997. Vol.~11. P.~341--359.
+
+\item \label{rfa:rulit:dual_annealing}
+Xiang~Y, Gubian~S, Suomela~B, Hoeng~J. Generalized Simulated Annealing for Efficient Global Optimization: the GenSA Package for R // The R Journal. 2013. Vol.~5, No.~1.
+
+\item \label{rfa:rulit:direct}
+Gablonsky~J., Kelley~C. A Locally-Biased form of the DIRECT Algorithm // Journal of Global Optimization. 2001. Vol.~21. P.~27--37.
+
+\item \label{rfa:rulit:basinhopping}
+Wales~D.J., Doye~J.P.K.
+Global Optimization by Basin-Hopping and the Lowest Energy Structures of Lennard-Jones Clusters Containing up to 110 Atoms // Journal of Physical Chemistry A.
+1997.
+Vol.~101.
+p.~5111.
+
+\item \label{rfa:rulit:shgo}
+Endres~S.C., Sandrock~C., Focke~W.W. A simplicial homology algorithm for lipschitz optimisation // Journal of Global Optimization. 2018.
+
+\item \label{rfa:rulit:nikitin2022automated}
+Filippou~K., Aifantis~G., Papakostas~G.A., Tsekouras~G.E.
+Structure learning and hyperparameter optimization using an automated machine learning ({AutoML}) pipeline // Information.
+2023.
+Vol.~14, No.
+4.
+p.~232.
+
+\item \label{rfa:rulit:MonthlyBeerArticle}
+Xu~N. Time Series Analysis on Monthly Beer Production in Australia~// Highlights in Science, Engineering and Technology. 2024. Vol.~94. P.~392--401. DOI: 10.54097/4z3krj13
+
+%10
+%\item \label{rfa:rulit:Kvasov2013}
+%Квасов~Д.Е., Сергеев~Я.Д. Методы липшицевой глобальной оптимизации в задачах управления~// Автоматика и телемеханика. 2013. №~9. C.~3--19.
+
+%17
+%\item \label{rfa:rulit:Sergeyev2020}
+%Sergeyev~Y.D., Candelieri~A., Kvasov~D.E., Perego~R.
+Safe global optimization of expensive noisy black-box functions in the $\delta$-Lipschitz framework~// Soft Comput.
+2020.
+Vol.~24, No.~23.
+P.~17715--17735.
+DOI: 10.1007/s00500-020-05030-3.
+
+%4
+%\item \label{rfa:rulit:Dongarra2022}
+%Dongarra~J.J. The evolution of mathematical software~// Commun. ACM. 2022. Vol.~65, No.~12. P.~66--72. DOI: 10.1145/3554977.
+
+%5
+%\item \label{rfa:rulit:Duwe2020}
+%Duwe~K., et al. State of the art and future trends in data reduction for high-performance computing~// Supercomput. Front. Innov. 2020. Vol.~7, No.~1. DOI: 10.14529/jsfi200101.
+
+%\item \label{rfa:rulit:fedot}
+%Automated modeling and machine learning framework {FEDOT} -- URL:  https://github.com/aimclub/FEDOT (дата обращения: 25.07.2025).
 

### 2025-07-28 2025/Problems of Informatics/article.tex
translate abstract

@@
 % аннотация англ.
-The article discusses  the problem of finding the best combination of hyperparameters of machine learning and artificial intelligence methods.
+The paper discusses  the problem of finding the best combination of hyperparameters of machine learning and artificial intelligence methods.
 In many cases, the efficiency (in a given metric) of the resulting solution for different values of hyperparameters can be quite different.
-В таких задачах актуальной является проблема некорректной работы исследуемых artificial intelligence (AI) and machine learning (ML) methods в некоторых (заранее неизвестных) подобластях области изменения гиперпараметров.
+In such problems, a significant issue is the potential for incorrect operation of the investigated artificial intelligence and machine learning methods within certain (a priori unknown) subregions of the hyperparameters search domain.
+Furthermore, the computational complexity of tuning makes manual or exhaustive search inappropriate.
+These characteristics have required the development of various intelligent automatic hyperparameter optimization methods.
+%В таких задачах актуальной является проблема некорректной работы исследуемых artificial intelligence (AI) and machine learning (ML) methods в некоторых (заранее неизвестных) подобластях области изменения гиперпараметров.
 При этом вычислительная сложность настройки делает невозможным использование ручного или полного перебора.
 Данные особенности определили необходимость разработки различных методов интеллектуального автоматического подбора гиперпараметров.
-С математической точки зрения такая задача может быть представлена как problem of finding a global minimum of a function, given in the form of a ``black box'' and computable only in some part of the search domain.
+From a mathematical point of view, such a task can be represented as the problem of finding a global minimum of a function, given in the form of a ``black box'' and computable only in some part of the search domain.
 In this case, each computation of the objective function value at some point of the feasible domain may require significant computing resources.
@@
 The existence of subdomains where the objective function is undefined can be interpreted as the existence of some hidden, a priori unknown constraints of the problem.
-The authors propose an approach to solving this type of problem, , which is an extension of the information-statistical global search algorithm (GSA) and takes into account the presence of undefined values of the objective function at some points.
-В рамках алгоритма происходит разбиение области поиска точками испытаний и оцениваются характеристики подобластей на основе значений целевой функции, вычисленных на их границах.
-В случае отсутствия информации о значении функции в точке в алгоритме используется предложенная авторами оценка, учитывающая размер исследуемой подобласти.
-To minimize the number of redundant trials in subdomains where the function is not defined, we used a special method parameter $\alpha$ that regulates the number of trial points in areas of non-computability.
+The authors propose an approach to solving this type of problem, which is an extension of the information-statistical global search algorithm (GSA) and takes into account the presence of undefined values of the objective function at some points.
+The algorithm partitions the search space with trial points and evaluates the characteristics of subregions based on the objective function values computed at their boundaries.
+If the function value at a point is unknown, the algorithm employs an estimate for this value, considering the size of the subregion under investigation.
+%В рамках алгоритма происходит разбиение области поиска точками испытаний и оцениваются характеристики подобластей на основе значений целевой функции, вычисленных на их границах.
+В случае отсутствия информации о значении функции в точке в алгоритме используется оценка этого значения, учитывающая размер исследуемой подобласти.
+To minimize the number of redundant trials in subdomains where the function is not defined, the method parameter $\alpha$ was used that regulates the number of trial points in regions of non-computability.
 The solution of multidimensional problems implemented through reducing them to one-dimensional optimization problems using space-filling curves (Peano curves).
-The article provides a detailed description and a diagram of the operation of the modified global search algorithm.
-The implementation of the global search algorithm for the case of a not-everywhere computable objective function (GSA-N) was based on the iOpt open source framework of intelligent optimization methods.
+The article provides a detailed description and a flowchart of the operation of the proposed search algorithm.
+The implementation of the global search algorithm for the case of a not everywhere computable objective function (GSA-N) was based on the iOpt open source framework of intelligent optimization methods.
 To carry out the experiments, a generator of test problems with hidden constraints GKLS-HC was developed.
@@
 The experimental results presented in the paper, obtained on a series of GKLS-HC test problems, demonstrate the reliability of global search and the efficiency of the developed algorithm.
-By adjusting the parameter $\alpha$, it was possible to achieve the same speed of GSA-N operation as the basic GSA.
+By adjusting the parameter $\alpha$, it was possible to achieve the same performance of GSA-N operation as the basic GSA.
 The paper also considered the behavior of global optimization algorithms of the scipy.optimize library when solving problems with non-computable domains.
@@
 The DIRECT and SHGO methods, although they solved the problems, were less effective than the developed GSA-N.
-Основным результатом являются эксперименты с модельными задачами настройки гиперпараметров, в которых возникают неопределенные значения оптимизируемой метрики качества.
-В этих задачах существуют недопустимые сочетания гиперпараметров, при которых настраиваемый метод возвращает бесконечное значение целевой функции.
-Так была успешно проведена настройка классического алгоритма классификации LinearSVC.
-АГП-Н успешно решил поставленную задачу, найдя лучшее сочетание гиперпараметров по сравнению со стандартными алгоритмами из библиотеки scikit-learn.
-Также был настроен метода предсказания значений временного ряда на базе фреймворка FEDOT.
-В ходе этого эксперимента также было проведено сравнение АГП-Н с алгоритмами настройки из другого широко известного фреймворка Optuna.
+Experiments were also conducted with hyperparameter tuning problems where undefined values of the quality metric arise.
+In these problems, certain hyperparameter combinations caused the method to return infinite values for the objective function.
+The LinearSVC classification algorithm was successfully tuned.
+GSA-N effectively solved the problem, identifying a better hyperparameter combination compared to standard scikit-learn algorithms.
+A time series prediction method from the FEDOT framework was also configured.
+During this experiment, GSA-N was compared to tuning algorithms from the popular Optuna framework.
+GSA-N achieved comparable performance in terms of the obtained the target metric value, significantly surpassing Optuna in solution time.
+%Также проведены эксперименты с модельными задачами настройки гиперпараметров, в которых возникают неопределенные значения оптимизируемой метрики качества.
+В этих задачах существовали сочетания гиперпараметров, при которых настраиваемый метод возвращает бесконечное значение целевой функции.
+Была успешно проведена настройка классического алгоритма классификации LinearSVC.
+GSA-N успешно решил поставленную задачу, найдя лучшее сочетание гиперпараметров по сравнению со стандартными алгоритмами из библиотеки scikit-learn.
+Также был настроен метод предсказания значений временного ряда из фреймворка FEDOT.
+В ходе этого эксперимента было проведено сравнение GSA-N с алгоритмами настройки из другого широко известного фреймворка Optuna.
 АГП-Н показал сравнимый результат (в терминах полученного значения целевой метрики), значительно обогнав Optuna по времени, затраченному на решение задачи.

### 2025-08-19 2025/Problems of Informatics/article.tex
last fixes

@@
 Данные особенности определили необходимость разработки различных методов интеллектуального автоматического подбора гиперпараметров.
-From a mathematical point of view, such a task can be represented as the problem of finding a global minimum of a function, given in the form of a ``black box'' and computable only in some part of the search domain.
+From a mathematical point of view, such a task can be re\-pre\-sen\-ted as the problem of finding a global minimum of a function, given in the form of a ``black box'' and computable only in some part of the search domain.
 In this case, each computation of the objective function value at some point of the feasible domain may require significant computing resources.
@@
 \item \label{rfa:enlit:Birect2020}
-Paulavi{\v c}ius~R., Sergeyev~Y.D., Kvasov~D.E., {\v Z}ilinskas~J. Globally-biased {BIRECT} algorithm with local accelerators for expensive global optimization~// 
+Paulavi{\v c}ius~R., Sergeyev~Y.D., Kvasov~D.E., {\v Z}ilinskas~J. Globally-biased {BIRECT} al\-go\-rithm with local accelerators for expensive global optimization~// 
 Expert Syst. Appl. 2020. Vol.~144. p.~113052. %DOI: 10.1016/j.eswa.2019.113052.
@@
 \item \label{rfa:enlit:Strongin2020}
-Strongin~R.G., Barkalov~K.A., Bevzuk~S.A. Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints~// 
+Strongin~R.G., Barkalov~K.A., Bevzuk~S.A. Global optimization method with dual Lip\-schitz constant estimates for problems with non-convex constraints~// 
 Soft Comput. 2020. Vol.~24, No.~16. P.~11853--11865. %DOI: 10.1007/s00500-020-05078-1.
@@
 2024.
-Vol.~1914.
+Vol.
+1914.
 P.~147--161.
@@
 Sysoev~A.V., Kozinov~E.A., Barkalov~K.A., Lebedev~I.G., Karchkov~D.A., Rodionov~D.M.
-Frejmvork metodov intellektualnoj evristicheskoj optimizacii iOpt // V kn.: Superkompyuternye dni v Rossii: Trudy mezhdunarodnoj konferencii.
+Frejmvork metodov intellektualnoj evristicheskoj optimizacii iOpt // V kn.: Su\-per\-kom\-pyu\-ter\-nye dni v Rossii: Trudy mezhdunarodnoj konferencii.
 2023.
@@
 \item \label{rfa:enlit:iOptGithub}
-Isxodnyj kod frejmvorka iOpt -- URL: https://github.com/aimclub/iOpt (data obrashheniya: 26.01.2025).
+Isxodnyj kod frejmvorka iOpt -- URL: https://github.com/aimclub/iOpt (data ob\-ra\-she\-ni\-ya: 26.01.2025).
 
 \item \label{rfa:enlit:iOptDocs}
-Dokumentaciya iOpt -- URL: https://iopt.readthedocs.io/ru/latest/ (data obrashheniya: 26.01.2025).
+Dokumentaciya iOpt -- URL: https://iopt.readthedocs.io/ru/latest/ (data ob\-ra\-she\-ni\-ya: 26.01.2025).
 
@@
 Filippou~K., Aifantis~G., Papakostas~G.A., Tsekouras~G.E.
-Structure learning and hyperparameter optimization using an automated machine learning ({AutoML}) pipeline // Information.
+Structure learning and hy\-per\-pa\-ra\-me\-ter optimization using an automated machine learning ({AutoML}) pipeline // Information.
 2023.
@@
 
+\item \label{rfa:enlit:FEDOT}
+Automated modeling and machine learning framework {FEDOT} -- URL:\\https://github.com/aimclub/FEDOT (data obrashheniya: 25.07.2025).
+
 \item \label{rfa:enlit:MonthlyBeerArticle}
-Xu~N. Time Series Analysis on Monthly Beer Production in Australia~// Highlights in Science, Engineering and Technology. 2024. Vol.~94. P.~392--401. DOI: 10.54097/4z3krj13
+Xu~N. Time Series Analysis on Monthly Beer Production in Australia~// Highlights in Science, Engineering and Technology. 2024. Vol.~94. P.~392--401.% DOI: 10.54097/4z3krj13
 
@@
 
-%\item \label{rfa:enlit:FEDOT}
-%Automated modeling and machine learning framework {FEDOT} -- URL:  https://github.com/aimclub/FEDOT (data obrashheniya: 25.07.2025).
+\item \label{rfa:renlit:optuna}
+Akiba T., Sano S., Yanase T., Ohta T., Koyama M.
+Optuna: A Next-Generation Hy\-per\-pa\-ra\-me\-ter Optimization Framework // In Proceedings: 25th ACM SIGKDD International Conference on Knowledge Discovery \& Data Mining.
+2019.
+P.
+2623--2631.%DOI: 10.1145/3292500.3330701
 
 \end{enumerate}
+
+%\newpage
 
@@
 В качестве второй модельной задачи рассмотрим задачу настройки гиперпараметров методов предсказания значений временного ряда (time series forecasting).
-Предсказание значений производилось с использованием фреймворка FEDOT [\ref{rfa:rulit:nikitin2022automated}]
+Предсказание значений производилось с использованием фреймворка FEDOT [\ref{rfa:rulit:nikitin2022automated}, \ref{rfa:rulit:fedot}]
 (фреймворк с открытым исходным кодом для задач автоматизированного моделирования и машинного обучения, AutoML).
@@
 
-Для сравнения поставленная задача также была решена при помощи известного фреймворка оптимизации гиперпараметров Optuna%\cite{OptunaURL,optuna}
-. Оба решателя были интегрированы в FEDOT и задействованы в вычислительном эксперименте.
+Для сравнения поставленная задача также была решена при помощи известного фреймворка оптимизации гиперпараметров Optuna [\ref{rfa:rulit:optuna}].
+Оба решателя были интегрированы в FEDOT и задействованы в вычислительном эксперименте.
 
@@
 \item \label{rfa:rulit:Frazier2018}
-Frazier~P.I. A Tutorial on Bayesian Optimization~// arXiv. 2018. DOI: 10.48550/arXiv.1807.02811.
+Frazier~P.I. A Tutorial on Bayesian Optimization~// arXiv. 2018.\\DOI: 10.48550/arXiv.1807.02811.
 
@@
 Paulavi{\v c}ius~R., Sergeyev~Y.D., Kvasov~D.E., {\v Z}ilinskas~J.
-Globally-biased {BIRECT} algorithm with local accelerators for expensive global optimization~// Expert Syst.
+Globally-biased {BIRECT} al\-go\-rithm with local accelerators for expensive global optimization~// Expert Syst.
 Appl.
@@
 \item \label{rfa:rulit:Audet2022}
-Audet~C., Batailly~A., Kojtych~S. Escaping unknown discontinuous regions in blackbox optimization~// SIAM J. Optim. 2022. Vol.~32, No.~3. P.~1843--1870. DOI: 10.1137/21m1420915.
+Audet~C., Batailly~A., Kojtych~S. Escaping unknown discontinuous regions in blackbox optimization~// SIAM J. Optim. 2022. Vol.~32, No.~3. P.~1843--1870. \\DOI: 10.1137/21m1420915.
 
@@
 Candelieri~A. Sequential model based optimization of partially defined functions
-under unknown constraints~// J. Glob. Optim. 2019. Vol.~79, No.~2. P.~281--303. DOI: 10.1007/s10898-019-00860-4.
+under unknown constraints~// J. Glob. Optim. 2019. Vol.~79, No.~2. P.~281--303. \\DOI: 10.1007/s10898-019-00860-4.
 
@@
 Usova~M.A., Barkalov~K.A. An Algorithm for Finding the Global Extremum of a Partially Defined Function~//
-Communications in Computer and Information Science. 2024. Vol.~1914. P.~147--161. DOI: 10.1007/978-3-031-52470-7{\_}13.
+Communications in Computer and Information Science. 2024. Vol. 1914. P.~147--161. DOI: 10.1007/978-3-031-52470-7{\_}13.
 
@@
 p.~232.
+
+\item \label{rfa:rulit:fedot}
+Automated modeling and machine learning framework {FEDOT} -- URL: \\https://github.com/aimclub/FEDOT (дата обращения: 25.07.2025).
 
@@
 
-%\item \label{rfa:rulit:fedot}
-%Automated modeling and machine learning framework {FEDOT} -- URL:  https://github.com/aimclub/FEDOT (дата обращения: 25.07.2025).
+\item \label{rfa:rulit:optuna}
+Akiba T., Sano S., Yanase T., Ohta T., Koyama M.
+Optuna: A Next-Generation Hyperparameter Optimization Framework // In Proceedings: 25th ACM SIGKDD International Conference on Knowledge Discovery \& Data Mining.
+2019.
+P.
+2623--2631.
+DOI: 10.1145/3292500.3330701
 

### 2025-08-19 2025/Problems of Informatics/article.tex
minor corrections

@@
 
-Основная сложность при решении  задач настройки гиперпараметров состоит  в  том, что поиск их оптимального сочетания требует для каждого выбранного варианта решения соответствующей исходной задачи машинного обучения, а значит, может быть довольно длительным по времени.
+Основная сложность настройки гиперпараметров состоит в том, что поиск их оптимального сочетания требует (для каждого выбранного варианта) решения исходной задачи машинного обучения, а значит, может быть довольно длительным по времени.
 Таким образом, любой подход будет ограничен в числе сочетаний гиперпараметров, которое метод глобальной оптимизации сможет проверить прежде, чем будет исчерпан доступный вычислительный ресурс.
@@
 
-Хорошие результаты, достигаемые методами байесовской оптимизации, основаны на априорном предположении, что целевая функция задачи соответствует определенной стохастической модели, например, гауссовскому процессу.
+Хорошие результаты при использовании методов байесовской оптимизации достигаются, когда целевая функция задачи соответствует определенной стохастической модели, например, гауссовскому процессу.
 Однако в глобальной оптимизации существуют и иные предположения о виде функции, которые дают замечательные результаты.
@@
 
-В рассматриваемых задачах настройки гиперпараметров актуальной становится новая проблема -- некорректная работа исследуемых алгоритмов ИИ и МО в некоторых (заранее неизвестных) подобластях области изменения гиперпараметров.
-При этом программные реализации настраиваемых методов могут вести себя абсолютно по-разному: в лучшем случае информируют пользователя о том, что задача не может быть решена, в худшем -- возвращают недопустимое значение (NaN, inf) в качестве найденного решения.
+В рассматриваемых задачах настройки гиперпараметров актуальной является проблема некорректной работы исследуемых алгоритмов ИИ и МО в некоторых (заранее неизвестных) подобластях области поиска. 
+При этом программные реализации настраиваемых методов могут вести себя абсолютно по-разному: в лучшем случае информируют исследователя о том, что задача не может быть решена, в худшем -- возвращают недопустимое значение (NaN, inf) в качестве найденного решения.
 Проблема возникновения недопустимых комбинаций, которая не имеет особого значения в случае <<ручной>> настройки, существенно осложняет работу фреймворков автоматической настройки гиперпараметров.
 
-Данное явление с математической точки зрения можно интерпретировать или как наличие в задаче некоторых скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], или как наличие неизвестных областей, в которых целевая функция не является непрерывной [\ref{rfa:rulit:Audet2022}], или как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Sergeyev2003},\ref{rfa:rulit:Strongin2020}].
+Указанное свойство с математической точки зрения можно интерпретировать либо как наличие в задаче некоторых скрытых ограничений [\ref{rfa:rulit:Stripinis2021}], либо как наличие неизвестных областей, в которых целевая функция не является непрерывной [\ref{rfa:rulit:Audet2022}], либо как частичную вычислимость целевой функции в области поиска [\ref{rfa:rulit:Candelieri2019},\ref{rfa:rulit:Sergeyev2003},\ref{rfa:rulit:Strongin2020}].
 В такой постановке задача оптимизации существенно усложняется, т.к.

### 2025-08-20 2025/Problems of Informatics/article.tex
last fix

@@
 \item \label{rfa:enlit:Sergeyev2003}
-Barkalov~K.A., Strongin~R.G. Metod globalnoj optimizacii s adaptivnym poryadkom proverki ogranichenij~// Zhurn. vy`chisl. matem. i matem. fiz. 2002. T.~42. No.~9. S.~1338--1350.
+Barkalov~K.A., Strongin~R.G. Metod globalnoj optimizacii s adaptivnym poryadkom proverki ogranichenij~// Zhurn. vy`chisl. matem. i matem. fiz. 2002. T.~42, No.~9. S.~1338--1350.
 
@@
 \item \label{rfa:rulit:MonthlyBeerArticle}
-Xu~N. Time Series Analysis on Monthly Beer Production in Australia~// Highlights in Science, Engineering and Technology. 2024. Vol.~94. P.~392--401. DOI: 10.54097/4z3krj13
+Xu~N. Time Series Analysis on Monthly Beer Production in Australia~// Highlights in Science, Engineering and Technology. 2024. Vol.~94. P.~392--401. DOI: 10.54097/4z3krj13.
 
@@
 2623--2631.
-DOI: 10.1145/3292500.3330701
+DOI: 10.1145/3292500.3330701.
 
@@
 Н.И.
-Лобачевского, e-mail: ilya.lebedev@itmm.unn.ru.
-Область научных интересов: методы глобальной и локальной оптимизации, параллельные алгоритмы, программирование для графических процессоров, CUDA.
+Лобачевского, Область научных интересов: методы глобальной и локальной оптимизации, параллельные алгоритмы, программирование для графических процессоров, CUDA.
 Является автором и соавтором 61 научной работы.
+E-mail: ilya.lebedev@itmm.unn.ru.
 \paragraph{Штанюк Антон Александрович} -- кандидат технических наук, доцент, доцент кафедры информатики и автоматизации научных исследований национального исследовательского нижегородского государственного университета им.
@@
 Лобачевского.
-E-mail: anton.shtanyuk@itmm.unn.ru.
 Количество печатных работ 53.
 Область научных интересов: моделирование и проектирование сложных систем, языки и технология программирования, обучение программированию, алгоритмы и структуры данных.
+E-mail: anton.shtanyuk@itmm.unn.ru.
 \paragraph{Баркалов Константин Александрович} -- 
@@
 Lobachevsky.
-e-mail: ilya.lebedev@itmm.unn.ru.
 His research interests include algorithms of global and local optimization, parallel algorithms, GPU programming, CUDA.
 He is the author or coauthor more than 30 papers in these areas.
+E-mail: ilya.lebedev@itmm.unn.ru.
 \paragraph{Anton Shtanyuk} -- PhD., Associate Professor at the Department of Informatics and Research Automation of the Institute of Information Technology, mathematics and mechanics at Lobachevsky State University of Nizhny Novgorod.
-E-mail: anton.shtanyuk@itmm.unn.ru The number of publications: 50.
+The number of publications: 50.
 Research interests: modelling and design of complex systems, programming languages and technology, programming teaching, algorithms and data structures.
+E-mail: anton.shtanyuk@itmm.unn.ru.
 \paragraph{Konstantin Barkalov} -- Doctor of Technical Sciences, Professor, Head of the Department of Mathematical Software and Supercomputer Technologies, Institute of Information Technologies, Mathematics and Mechanics, Lobachevsky State University of Nizhny Novgorod.

### 2025-09-13 2025/MMST_DecisionTree/ParallelDecisionTree.tex
translate up the Section 3

@@
 %
-\title{Параллельный запуск локальных методов алгоритме глобального поиска с использованием деревьев решений}
+\title{Parallel global optimization algorithm employing decision trees for launching local methods}
 %
@@
 %
-\author{К.А. Баркалов\inst{1}\orcidID{0000-1111-2222-3333} \and
-И.Г. Лебедев\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Я.В. Селенко\inst{3}\orcidID{2222--3333-4444-5555}}
+\author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
+Ilya Lebedev\orcidID{0000-0002-8736-0652} \and
+Dmitry Silenko}
 %
-\authorrunning{F. Author et al.}
+\authorrunning{K. Barkalov et al.}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Нижегородский государственный университет им. Н.И. Лобачевского, Нижний Новгород, Российская Федерация }
+\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia}
 %
@@
 \begin{abstract}
-В данной работе предлагается параллельный подход к запуску локальных методов при решении задачи глобальной оптимизации алгоритмом глобального поиска.
-Целевая функция задачи задана как черный ящики и от нее требуется лишь удовлетворять условию Липшица с априори неизвестной константой.
+%В данной работе разработан подход к параллельному запуску локальных методов при решении задачи глобальной оптимизации с помощью алгоритма глобального поиска.
+Целевая функция задачи задана как черный ящик; предполагается, что она удовлетворяет условию Липшица с априори неизвестной константой.
 В работе рассматривается метод выбора окрестности локальных экстремумов целевой функции на основе анализа накопленной поисковой информации.
-Проведение такого анализа с использованием методов машинного обучения позволяет принять решение о параллельном запуске локальных методов, что может ускорить сходимость алгоритма.
-Это предположение было подтверждено результатами численных экспериментов, демонстрирующих ускорение при решении ряда тестовых задач.
+Проведение такого анализа с использованием методов машинного обучения позволяет принять решение о параллельном запуске локальных методов, что позволяет ускорить сходимость алгоритма.
+Это предположение было подтверждено результатами численных экспериментов, демонстрирующих ускорение при решении серии тестовых задач.
+
+This paper presents an approach for the parallel launch of local optimization methods within a global search algorithm framework, aimed at solving global optimization problems.
+The objective function is treated as a black-box and is assumed that it satisfies the Lipschitz condition with an unknown constant.
+The proposed method focuses on selecting the region of attraction of local extrema of the objective function based on the analysis of accumulated search data.
+Employing machine learning techniques for this analysis enables informed decisions regarding the parallel execution of local methods, leading to improved algorithm convergence.
+The effectiveness of this approach is validated through numerical experiments, which demonstrate accelerated performance on a set of benchmark problems.
 
 \keywords{
-	Глобальная оптимизация $\cdot$ 
-	Многоэкстремальные функции $\cdot$ 
-	Параллельные вычисления $\cdot$ 
-	Машинное обучение $\cdot$ 
-	Древо решений $\cdot$}
+Global optimization $\cdot$
+Multi-extremal functions $\cdot$
+Parallel computing $\cdot$
+Machine learning $\cdot$
+Decision tree $\cdot$}
 \end{abstract}
@@
 
-\section{Введение}
+\section{Introduction}
 
-В настоящей работе рассматриваются параллельные алгоритмы решения многомерных задач глобальной оптимизации.
+%В настоящей работе рассматриваются параллельные алгоритмы решения многомерных задач глобальной оптимизации.
 Такие задачи часто возникают в тех случаях, когда необходимо подобрать значения параметров исследуемой математической модели, при которых результаты моделирования лучше всего соответствуют экспериментальным данным.
 Для решения задач этого класса известно множество алгоритмов: от метаэвристических, основанных на идее случайного поиска \cite{Ferreiro2013,Garcia2014,Langdon2011}, до детерминированных алгоритмов, гарантирующих сходимость к глобальному минимуму \cite{Evtushenko2009,He2008,Paulavicius2011}.
+This paper investigates parallel algorithms for solving multidimensional global optimization problems These problems commonly occur when optimizing the parameters of a mathematical model to achieve the best possible agreement between simulation results and experimental observations.
+Numerous algorithms have been developed to address this class of problems, varying from metaheuristic techniques based on the idea of random search \cite{Ferreiro2013,Garcia2014,Langdon2011} to deterministic algorithms designed to guarantee convergence to a global minimum \cite{Evtushenko2009,He2008,Paulavicius2011}.
 
-Поскольку в реальных задачах глобальной оптимизации каждое вычисление значения функции (далее \textit{испытание}) является операцией, требующей больших вычислительных затрат, количество таких попыток приходится сокращать.
+%Поскольку в реальных задачах глобальной оптимизации каждое вычисление значения функции (далее \textit{trial}) является операцией, требующей больших вычислительных затрат, количество таких попыток приходится сокращать.
 Это может быть достигнуто за счет преднамеренного выбора вариантов в ходе поиска оптимального решения, отсекая бесперспективные поисковые под области и исследуя только те, в которых можно найти решение задачи.
@@
 Решение о запуске локального метода будет приниматься с использованием дерева решений.
+In real-world global optimization problems, each function evaluation (hereafter referred to as a \textit{trial}) is a computationally expensive operation, making it necessary to reduce the number of such trials.
+This may be achieved by the purposeful selection of alternatives during the course of optimization, rejecting unpromising search sub-regions and exploring only those where a solution can be found.
+The Global Search Algorithm (GSA) is based on this idea \cite{Strongin2000}.
+In this work, we attempted to combine the GSA and a local optimization method \cite{Nocedal, Kelley} (the Hooke-Jeeves pattern search method \cite{HookJeeves}) to reduce the number of trials performed.
+The decision to initiate the local method will be made using a decision tree.
 
-\section{Алгоритмы}\label{SecA}
+\section{Problem statement and solution strategies}\label{SecA}
 
-Рассмотрим задачу поиска глобального минимума функции $\varphi(y)$ на гиперинтервале $D=\{ y\in\ R^N:\ a_i\le\ y_i\le\ b_i,\ 1\le\ i\le\ n \}$. 
+We consider the problem of finding the global minimum of the function $\varphi(y)$ over a hyperinterval $D$, 
 \begin{equation} \label{sec:problem}   
@@
 \end{equation}
-где $a,b \in R$ — заданные векторы.
+where $a,b \in R$ are given vectors.
 
-Также предположим, что функция удовлетворяет условию Липшица с \textit{априори} неизвестной константой $L$, что соответствует ограниченному изменению значений функции при ограниченном изменении аргумента.
+We assume that the function satisfies the Lipschitz condition 
 \begin{displaymath} 
-	|\varphi(y_1)-\varphi(y_2)|\leq L\parallel y_1-y_2 \parallel ,y_1,y_2 \in D, 0<L< \infty. 
+	|\varphi(y_1)-\varphi(y_2)|\leq L\parallel y_1-y_2 \parallel ,y_1,y_2 \in D, 0<L< \infty, 
 \end{displaymath}
-Это предположение можно рассматривать (применительно к реальным задачам) как отражение ограниченной мощности, порождающей изменения в моделируемой системе.
+with an a priori unknown constant $L$, which corresponds to a bounded change in the function's values for a bounded change in the argument.
+This assumption can be viewed (in the context of real-world problems) as reflecting the limited power generating changes in the simulated system.
 
-Численное решение задачи (\ref{sec:problem}) сводится к построению оценки $y_k^\ast\in D$, соответствующей некоторому понятию близости к точке $y^\ast$ (например, ${ ||y^\ast-y}_k^\ast||\le\ \varepsilon$, где $\varepsilon\geq0$ — заданная точность) на основе конечного числа $k$ вычисленных значений целевой функции.
-Применительно к классу рассматриваемых задач предполагается, что целевая функция $\varphi(y)$ может быть определена алгоритмически, в результате выполнения некоторой подпрограммы или библиотеки.
+The numerical solution of the problem (\ref{sec:problem}) reduces to constructing an estimate $y_k^\ast\in D$ that corresponds to some notion of proximity to a point $y^\ast$ (e.g., ${ ||y^\ast-y}_k^\ast||\le\ \varepsilon$, where $\varepsilon\geq0$ is a given accuracy) based on a finite number $k$ of computed values of the objective function.
+With respect to the class of problems considered, we assume that the objective function $\varphi(y)$ can be algorithmically defined through the execution of a subroutine.
 
-При решении многомерных задач применяется уменьшение размерности (т.е.
-сведение многомерной задачи к эквивалентной одномерной) с помощью кривых Пеано.
-Они позволяют свести многомерную задачу оптимизации в области $D$ к одномерной задаче минимизации в интервале $[0, 1]$.
+When solving multidimensional problems, dimensionality reduction is employed (i.e., reducing the multidimensional problem to an equivalent one-dimensional one) using Peano curves.
+These curves allow transforming the multidimensional optimization problem in the region $D$ into a one-dimensional minimization problem on the interval [0, 1].
 \begin{displaymath}
@@
 \end{displaymath}
-где функция $\varphi(y(x^\ast))$ удовлетворяет более общему условию Гельдера
+where function $\varphi(y(x^\ast))$ satisfies the more general Hölder condition
 \begin{displaymath}
@@
 
-Поэтому вместо исходной задачи минимизации функции $\varphi(y)$ в области $D$ можно рассмотреть минимизацию одномерной функции $f(x)=\varphi(y(x)) $, удовлетворяющий условию Гельдера для $ x\in [0,1]$.
+Consequently, rather than addressing the original problem of minimizing the function $\varphi(y)$ within the domain $D$, we can consider minimizing the one-dimensional function $f(x)=\varphi(y(x))$, subject to the Hölder condition for $x \in [0, 1]$.
 
-Чтобы определить, в какой момент лучше запускать локальный метод в рамках глобального поиска, воспользуемся алгоритмом, основанным на дереве решений.
+%Чтобы определить, в какой момент лучше запускать локальный метод в рамках глобального поиска, воспользуемся алгоритмом, основанным на дереве решений.
 На первом этапе поиска происходит накопление поисковой информации в виде результатов выполненных испытаний.
 На основе результатов испытаний проводилось обучение дерева решений, позволяющего получать кусочно-постоянную аппроксимацию целевой функции и на ее основе прогнозировать поведение целевой функции.
+To identify when to best launch a local method  within the framework of a global search, we employ an algorithm based on a decision tree.
+The first stage of the search involves accumulating search information in the form of results from performed trials.
+Based on these trial results, a decision tree is trained, allowing us to obtain a piecewise-constant approximation of the objective function and, consequently, to predict the behavior of the objective function.
 
-Дерево решений — это инструмент для автоматизированного анализа больших массивов данных, применяемый в машинном обучении.
+%Дерево решений — это инструмент для автоматизированного анализа больших массивов данных, применяемый в машинном обучении.
 Его можно использовать как для решения задач классификации, так и для регрессии.
@@
 Обозначим значение, вычисленное деревом решений в точке $y$, как $z' = \psi(y)$.
-На рис. \ref{fig:fig2} представлен график целевой функции $\varphi(y)$; На рис. \ref{fig:fig2_2} представлена соответствующая аппроксимация $\psi(y)$, построенная с помощью дерева решений.
+A decision tree is a tool for automated analysis of large datasets used in machine learning.
+It can be used for both classification and regression problems.
+In the case of regression, each leaf of the tree is assigned a constant.
+Therefore, the resulting approximating function is piecewise-constant.
+In our case, the decision tree constructs a function $\psi(y)$ as a piecewise-constant approximation of $\varphi(y)$ in the search space.
+Let $z' = \psi(y)$ denote the value computed by the decision tree at point $y$.
 
+Fig. \ref{fig:fig2} shows the graph of the objective function $\varphi(y)$; Fig. \ref{fig:fig2_2} presents the corresponding approximation $\psi(y)$ constructed using a decision tree.
 \begin{figure}
@@
 			\includegraphics[width=1\linewidth]{figure/fig5.png}
-			\caption{Целевая функция $\varphi(y)$ из тестового класса} %% подпись к рисунку
+			\caption{Graph of the benchmark objective function $\varphi(y)$} %% подпись к рисунку
 			\label{fig:fig2}
@@
 			\includegraphics[width=1\linewidth]{figure/fig4.png}
-			\caption{Кусочно-постоянная аппроксимация $\psi(y)$ функции $\varphi(y)$} %% подпись к рисунку
+			\caption{Graph of the piecewise-constant approximation $\psi(y)$ of the function $\varphi(y)$} %% подпись к рисунку
 			\label{fig:fig2_2}
@@
 
-При реализации алгоритма поиска областей притяжения локальных минимумов для построения дерева решений использовались алгоритмы из библиотеки OpenCV.
+%При реализации алгоритма поиска областей притяжения локальных минимумов для построения дерева решений использовались алгоритмы из библиотеки OpenCV.
 OpenCV — это библиотека с открытым исходным кодом алгоритмов компьютерного зрения, обработки изображений и численных алгоритмов общего назначения.
 Более подробную информацию о дереве решений можно найти в \cite{Brahmbhatt2013}.
+To implement the algorithm for identifying the regions of attraction for local minima to construct the decision tree algorithms from the OpenCV library were employed.
+OpenCV is an open-source library encompassing computer vision algorithms, image processing methods, and general-purpose numerical routines.
+Detailed information on decision trees may be found in \cite{Brahmbhatt2013}.
 
-После определения с помощью дерева решений области притяжения локального минимума, из ближайшей точки запускаем локальный метод.
-Алгоритм поиска локального минимума Хука--Дживса относится к классу методов нулевого порядка.
+%После определения с помощью дерева решений области притяжения локального минимума, из точки ближайшего испытания, проведенного в этой области, запускается локальный метод.
+Мы использовали метод Хука--Дживса, который относится к классу методов нулевого порядка.
 Его правила вычисления представляют собой комбинацию исследовательского поиска (для выбора направления) и поиска в выбранном направлении \cite{Himmelblau72}.
+Having identified a basin of attraction for a local minimum using the decision tree, a local method is launched from the nearest trial point conducted within that basin.
+We employed the Hooke-Jeeves method, classified as a zero-order technique.
+Its computational rules combine an exploratory search (for direction selection) with a search along the selected direction \cite{Himmelblau72}.
 
@@
 			\includegraphics[width=1\linewidth]{figure/fig1.pdf} 
-			\caption{Пример итераций алгоритма Хука-Дживса} %%  подпись к рисунку 
+			\caption{An example showing the iterations of the Hooke-Jeeves method} %%  подпись к рисунку 
 			\label{fig:fig1} 
@@
 
-На рис.
-\ref{fig:fig1} показан пример работы алгоритма.
-Отображаются линии  уровня, закрашенные кружки обозначают успешные шаги, незакрашенные кружки соответствуют неудачным шагам, выполненным во время исследовательского поиска.
+Fig.
+\ref{fig:fig1} visually demonstrates the local algorithm in action.
+Level lines are shown with filled circles representing successful steps and open circles indicating unsuccessful steps from the exploratory search.
 

### 2025-09-13 2025/MMST_DecisionTree/ParallelDecisionTree.tex
Initial part of the Experiments section is translated

@@
 
-\section{Многомерный параллельный алгоритм глобального поиска}\label{SecGSA}
+\section{Global search algorithm with local methods starting in parallel}\label{SecGSA}
 
-На начальном этапе мастер-процесс (пусть он имеет идентификатор $0$) параллельно запускает $p$ испытаний в $p$ разных точках $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$ отрезка $[0,1]$.
-Прообразами двух точек являются граничные $x^1=0,x^p=1$; остальные — внутренние точки $x^i\in\left(0,1\right),i=2,\ldots,p-1$.
+%TODO
+%Здесь нужны общие слова про параллельную программу с несколькими процессами, мастером и рабочими
 
-Теперь предположим, что алгоритм уже сделал $K\geq0$ итераций и проведено $k$ испытаний.
+At the initial stage, the master process (let it have the identifier 0) launches $p$ trials in parallel at $p$ different points $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$, where preimages $\{x^1,x^2,\ldots,x^p\}$ are form the interval $[0,1]$.
+Two preimages are the boundary points $x^1=0, x^p=1$; the remaining preimages are interior points $x^i\in\left(0,1\right),i=2,\ldots,p-1$.
+
+Let us now assume that the algorithm has completed $K \geq 0$ iterations and $k \geq K$ trials have been performed.
+
+%TODO
+%Здесь нужно связующее предложение
 
@@
 % переводить до этого раздела
-\section{Численные эксперименты}\label{SecR}
+\section{Numerical experiments}\label{SecR}
 
-Численные эксперименты проводились на суперкомпьютере Лобачевского. Каждый узел имел по два процессора Intel Sandy Bridge E5-2660 2,2 ГГц, 64 Гб оперативной памяти.
+%Численные эксперименты проводились на суперкомпьютере Лобачевский. Каждый узел имел по два процессора Intel Sandy Bridge E5-2660 2,2 ГГц, 64 Гб оперативной памяти.
+Numerical experiments were conducted on the Lobachevsky supercomputer. Each supercomputer node had two Intel Sandy Bridge E5-2660 2.2 GHz processors and 64 GB of RAM.
 
-В экспериментах использовался генератор тестовых задач GKLS, который может генерировать задачи многоэкстремальной оптимизации с известными свойствами: точка глобального минимума, количество локальных минимумов и др.
+%В экспериментах использовался генератор тестовых задач GKLS, который может генерировать задачи многоэкстремальной оптимизации с известными свойствами: точка глобального минимума, количество локальных минимумов и др.
+The GKLS test problem generator was used in the experiments.
+This generator can create multi-extremal optimization problems with known properties, such as the location of the global minimum, the number of local minima, etc.
 
-Ниже представлены результаты работы синхронного параллельного алгоритма глобального поиска  с использованием дерева решений для нахождения областей притяжения локальных минимумов.
+%Ниже представлены результаты работы синхронного параллельного алгоритма глобального поиска  с использованием дерева решений для нахождения областей притяжения локальных минимумов.
 Численное сравнение проводилось на двух классах функций GKLS (Simple и Hard из \cite{Sergeyev2006}) размерностей 2, 3, 4 и 5.
@@
 
-Критерием остановки алгоритмов было попадание очередных пробных точек в $\varepsilon$-близость к истинному глобальному минимуму.
+The results below show the performance of a synchronous parallel global search algorithm using a decision tree to find basins of attraction for local minima.
+The numerical comparison was performed on two classes of GKLS functions (Simple and Hard from \cite{Sergeyev2006}) with dimensions 2, 3, 4, and 5.
+These two classes differ in the size of the basin of attraction for the global minimum.
+The simple class of problems has a radius of attraction that is three times larger than that of the hard class.
+
+Критерием остановки алгоритмов было попадание очередных точек испытаний в $\varepsilon$-окрестность истинного глобального минимума.
 В таблицах \ref{tab:1} \ref{tab:2} сравниваются усредненные числа итераций, выполняемых алгоритмами при решении задач, и значения ускорения времени относительно последовательного прогона.

### 2025-09-14 2025/MMST_DecisionTree/ParallelDecisionTree.tex
Добавлен текст

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Классические алгоритмы локальной оптимизации предназначены для нахождения единственного экстремума (локального минимума или максимума) в окрестности начального приближения.
+В зависимости от используемой информации о целевой функции, эти методы принято классифицировать на:
+\begin{itemize}
+	\item \textbf{Методы нулевого порядка}, которые используют только значения функции.
+	\item \textbf{Градиентные методы (первого порядка)}, требующие вычисления первых производных.
+	\item \textbf{Методы второго порядка} (например, метод Ньютона), которые, помимо прочего, требуют вычисления матрицы вторых производных (гессиана).
+\end{itemize}
+
+Важно подчеркнуть, что применение градиентных методов (первого и, в особенности, второго порядка) для оптимизации функций типа черного ящика сопряжено с фундаментальными трудностями.
+Основная проблема заключается в отсутствии аналитического описания функции, что делает невозможным прямое вычисление производных и приводит к необходимости их численной оценки, что, в свою очередь, может быть вычислительно затратно и вносить дополнительные погрешности \cite{Kelley}.
+В связи с этим, в рамках данного исследования дальнейшее рассмотрение будет ограничено исключительно методами нулевого порядка, не требующими вычисления производных.
+
+Ключевой вопрос при гибридизации глобального и локального поиска заключается в определении оптимального момента для запуска локального метода.
+Для решения этой задачи предлагается использовать адаптивный алгоритм, основанный на машином обучении, а именно на построении дерева решений.
+
+На первоначальном этапе глобального поиска информация о поведении целевой функции накапливается в виде набора вычисленных значений в точках испытаний.
+Эти данные используются для обучения модели дерева решений.
+Обученная модель строит кусочно-постоянную аппроксимацию целевой функции, которая позволяет прогнозировать её поведение в ещё не исследованных областях.
+
+Механизм принятия решения о запуске локального поиска заключается в следующем: для точки, подозрительной на наличие в её окрестности локального минимума, с помощью аппроксимирующей модели анализируются значения в соседних точках.
+На основе этого анализа делается вывод о том, попадает ли следующая точка испытания глобального алгоритма в область притяжения данного локального минимума.
+Это позволяет своевременно запустить локальную оптимизацию для эффективного уточнения решения в данной области.
+
+Логика изложения материала в данной статье организована следующим образом:
+\begin{itemize}
+	\item \textbf{Раздел \ref{SecA}} посвящен описанию методологической базы исследования.
+В нем излагается основная идея используемого алгоритма глобального поиска и детализируются вычислительные правила его работы.
+Кроме того, раздел включает описание применяемого метода локальной оптимизации (нулевого порядка) и обосновывает методику построения аппроксимации функции с помощью деревьев решений.
+	\item \textbf{Раздел \ref{SecGSA}} представляет ключевой вклад работы — новый гибридный алгоритм, который интегрирует глобальный поиск с запуском локальных процедур, управляемым на основе прогноза от модели дерева решений.
+	\item \textbf{Раздел \ref{SecR}} содержит результаты вычислительных экспериментов, проведенных на параллельной вычислительной системе, их анализ и интерпретацию, а также формулировку основных заключительных выводов.
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Problem statement and solution strategies}\label{SecA}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Данный подход рассматривается в работах \cite{Sergeyev2022, Usova2024}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %Чтобы определить, в какой момент лучше запускать локальный метод в рамках глобального поиска, воспользуемся алгоритмом, основанным на дереве решений.
@@
 На основе результатов испытаний проводилось обучение дерева решений, позволяющего получать кусочно-постоянную аппроксимацию целевой функции и на ее основе прогнозировать поведение целевой функции.
-To identify when to best launch a local method  within the framework of a global search, we employ an algorithm based on a decision tree.
+To identify when to best launch a local method  within the framework of a global search, we employ an algorithm based on a decision tree  \cite{Barkalov2023_2}.
 The first stage of the search involves accumulating search information in the form of results from performed trials.
@@
 Обозначим значение, вычисленное деревом решений в точке $y$, как $z' = \psi(y)$.
-A decision tree is a tool for automated analysis of large datasets used in machine learning.
+%A decision tree is a tool for automated analysis of large datasets used in machine learning.
 It can be used for both classification and regression problems.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%Дерево решений представляет собой широко распространенный и мощный инструмент машинного обучения, предназначенный для автоматического выявления сложных зависимостей в больших многомерных массивах данных.
+В контексте нашего исследования данный метод применяется для построения вспомогательной аппроксимирующей модели с целью ускорения процесса глобальной оптимизации.
+
+%С формальной точки зрения, используемая модель представляет собой бинарное дерево, то есть такую древовидную структуру, где каждый внутренний (нелистовой) узел имеет ровно два дочерних узла.
+Этот универсальный инструмент может применяться для решения двух типов задач: классификации и регрессии.
+В задачах классификации результатом обработки объекта является отнесение его к одному из классов, для чего каждый терминальный узел (лист) дерева помечается соответствующей меткой класса, при этом несколько листьев могут содержать идентичные метки.
+В задачах регрессии, что более релевантно для нашего случая, каждому листу ставится в соответствие некоторое постоянное (скалярное) значение.
+Таким образом, итоговая регрессионная модель представляет собой кусочно-постоянную функцию, принимающую на различных участках области определения фиксированные значения.
+
+%В рамках предложенного подхода дерево решений используется для построения вспомогательной функции $\psi(y)$, которая служит кусочно-постоянной аппроксимацией некоторой заданной функции $\varphi(y)$ в пределах исследуемой области поиска.
+Значение, вычисляемое данной моделью в произвольной точке $y$, обозначается как $z' = \psi(y)$.
+Эта аппроксимация позволяет эффективно оценивать поведение целевой функции и принимать решения о дальнейшем направлении вычислительных ресурсов в процессе оптимизации.
+
+A decision tree is a widespread and powerful machine learning tool designed to automatically identify complex dependencies in large multidimensional datasets.
+In the context of our research, this method is used to build an auxiliary approximating model in order to accelerate the global optimization process.
+
+From a formal point of view, the model used is a binary tree, that is, a tree structure where each internal (non-leaf) node has exactly two child nodes.
+This versatile tool can be used to solve two types of problems: classification and regression.
+In classification tasks, the result of processing an object is to assign it to one of the classes, for which each terminal node (leaf) of the tree is marked with the corresponding class label, while several leaves may contain identical labels.
+In regression problems, which is more relevant for our case, each sheet is assigned a certain constant (scalar) value.
+Thus, the final regression model is a piecewise constant function that takes fixed values in different parts of the definition area.
+
+In the framework of the proposed approach, a decision tree is used to construct an auxiliary function $\psi(y)$, which serves as a piecewise constant approximation of a given function $\varphi(y)$ within the search area under study.
+The value calculated by this model at any point $y$ is denoted as $z' = \psi(y)$.
+This approximation makes it possible to effectively evaluate the behavior of the objective function and make decisions about the further direction of computing resources in the optimization process.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Fig. \ref{fig:fig2} shows the graph of the objective function $\varphi(y)$; Fig. \ref{fig:fig2_2} presents the corresponding approximation $\psi(y)$ constructed using a decision tree.
@@
 	\begin{center}
-		\begin{minipage}[h]{0.7\linewidth}
+		\begin{minipage}[h]{0.8\linewidth}
 			\includegraphics[width=1\linewidth]{figure/fig5.png}
@@
 	\begin{center}
-		\begin{minipage}[h]{0.7\linewidth}
+		\begin{minipage}[h]{0.8\linewidth}
 			\includegraphics[width=1\linewidth]{figure/fig4.png}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Исследующий поиск определяется следующим образом: 
+\begin{itemize}[label=$\bullet$] 
+	\item Определяется значение шага (он различен для каждого направления координат и может измениться во время поиска). 
+	\item Шаг поиска считается успешным, если значение целевой функции в контрольной точке не превышает значение исходной целевой функции. 
+	\item В противном случае нужно вернуться к предыдущему пункту и сделать шаг в обратном направлении. 
+	\item После перебора всех N координат исследовательский поиск заканчивается. Полученная точка называется базовой точкой.
+\end{itemize}
+Поиск по образцу состоит из выполнения одного шага от полученной базовой точки вдоль линии, соединяющей ее с предыдущей базовой точкой.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{figure} 
 	\begin{center} 
-		\begin{minipage}[h]{0.8\linewidth} 
+		\begin{minipage}[h]{1.0\linewidth} 
 			\includegraphics[width=1\linewidth]{figure/fig1.pdf} 
@@
 в \cite{Strongin2000,Barkalov2016}.
-Модификации с учетом наличия ограничений-неравенств, а также информация о производной целевой функции представлены в \cite{Barkalov2002,Gergel1997}.
+Модификации с учетом наличия ограничений-неравенств, а также информация о производной целевой функции представлены в \cite{Barkalov2002, Gergel1997, Barkalov2023, Gegrel2021}.
 
@@
 
-Критерием остановки алгоритмов было попадание очередных точек испытаний в $\varepsilon$-окрестность истинного глобального минимума.
-В таблицах \ref{tab:1} \ref{tab:2} сравниваются усредненные числа итераций, выполняемых алгоритмами при решении задач, и значения ускорения времени относительно последовательного прогона.
-Распараллеливание выполнено по технологии MPI, эксперимент выполнен на 8, 16, 32 и 64 процессах.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Критерием остановки алгоритмов было попадание очередных точек испытания в $\varepsilon$-близость к истинному глобальному минимуму.
+В таблицах \ref{tab:1} и \ref{tab:2} приведено среднее время решения и ускорение по времени относительно последовательного запуска, выполняемых алгоритмами при решении задач оптимизации.
+На рис.
+\ref{fig:fig4} приведено ускорение по числу итераций.
+Распараллеливание выполнено с использованием технологии MPI, эксперименты проводились на 1, 8, 16, 32 и 64 процессах.
 
 \begin{table}[!ht]
-	\caption{Ускорение, выполненных разными алгоритмами}
+	\caption{Среднее время решения}
+	\label{tab:1}
+	\centering	
+	\begin{tabular}{|l|l|l|l|l|l|}
+		\hline
+		2 & 1 mpi & 8 mpi & 16 mpi & 32 mpi & 64 mpi  \\ \hline
+		hard & 12,71 & 1,98 & 1,71 & 0,87 & 1,65  \\ \hline
+		simple & 6,35 & 0,57 & 1,27 & 0,95 & 0,21  \\ \hline
+		3 & ~ & ~ & ~ & ~ &  \\ \hline
+		hard & 14,19 & 5,59 & 2,10 & 0,75 & 0,60  \\ \hline
+		simple & 5,92 & 1,63 & 1,17 & 0,38 & 0,41  \\ \hline
+		4 & ~ & ~ & ~ & ~ &   \\ \hline
+		hard & 66,29 & 34,93 & 8,53 & 4,95 & 2,93  \\ \hline
+		simple & 31,88 & 9,44 & 4,71 & 4,67 & 5,51  \\ \hline
+		5 & ~ & ~ & ~ & ~ &  \\ \hline
+		hard & 324,04 & 52,42 & 17,55 & 11,48 & 10,60 \\ \hline
+		simple & 53,77 & 6,62 & 3,38 & 1,99 & 1,70  \\ \hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[!ht]
+	\caption{Ускорение по времени}
 	\label{tab:2}
@@
 		2 & 8 mpi & 16 mpi & 32 mpi & 64 mpi  \\ \hline
-		hard & 6,416962112 & 7,429653192 & 16,44977408 & 7,698208433  \\ \hline
-		simple & 11,19015498 & 5,017982941 & 13,58978911 & 30,41682225  \\ \hline
+		hard & 6,42 & 7,43 & 16,45 & 7,70  \\ \hline
+		simple & 11,19 & 5,018 & 13,59 & 30,42  \\ \hline
 		3 & ~ & ~ & ~ &   \\ \hline
-		hard & 4,010618178 & 6,760486061 & 22,40788476 & 23,48393504  \\ \hline
-		simple & 3,975141387 & 5,069355695 & 14,97196995 & 14,48568142  \\ \hline
+		hard & 4,01 & 6,76 & 22,41 & 23,48  \\ \hline
+		simple & 3,98 & 5,07 & 14,97 & 14,49  \\ \hline
 		4 & ~ & ~ & ~ &   \\ \hline
-		hard & 3,516700024 & 7,767619059 & 15,63797476 & 22,60877765  \\ \hline
-		simple & 4,204314065 & 6,771626114 & 7,265190222 & 5,789337963  \\ \hline
+		hard & 3,52 & 7,77 & 15,64 & 22,61  \\ \hline
+		simple & 4,20 & 6,77 & 7,27 & 5,79  \\ \hline
 		5 & ~ & ~ & ~ &   \\ \hline
-		hard & 10,28406793 & 18,45856679 & 38,75628766 & 30,57030882  \\ \hline
-		simple & 10,79552171 & 15,93214599 & 31,96240356 & 31,65420434  \\ \hline
+		hard & 10,28 & 18,46 & 38,76 & 30,57  \\ \hline
+		simple & 10,80 & 15,93 & 31,96 & 31,65  \\ \hline
 	\end{tabular}
@@
 
-Как видно из таблицы \ref{tab:2}, ускорение итерации достаточно велико для задач любой размерности.
+%\begin{table}[!ht]
+%	\caption{Ускорениепо итерациям}
+%	\label{tab:3}
+%	\centering	
+%	\begin{tabular}{|l|l|l|l|l|}
+	%		\hline
+	%		2 & 8 mpi & 16 mpi & 32 mpi & 64 mpi  \\ \hline
+	%		hard & 37,61 & 49,11 & 50,04 & 34,58  \\ \hline
+	%		simple & 52,06 & 19,72 & 63,02 & 307,59  \\ \hline
+	%		3 & ~ & ~ & ~ &   \\ \hline
+	%		hard & 15,03 & 92,04 & 47,28 & 89,74  \\ \hline
+	%		simple & 10,25 & 44,70 & 44,29 & 76,87  \\ \hline
+	%		4 & ~ & ~ & ~ &   \\ \hline
+	%		hard & 38,14 & 106,64 & 58,40 & 117,56  \\ \hline
+	%		simple & 30,75 & 79,84 & 29,41 & 40,43  \\ \hline
+	%		5 & ~ & ~ & ~ &   \\ \hline
+	%		hard & 162,66 & 269,09 & 201,95 & 206,94  \\ \hline
+	%		simple & 46,13 & 176,70 & 98,42 & 191,00  \\ \hline
+	%	\end{tabular}
+%\end{table}
+
+\begin{figure} 
+	\begin{center} 
+		\begin{minipage}[h]{0.9\linewidth} 
+			\includegraphics[width=1\linewidth]{figure/fig6.pdf} 
+			\caption{Ускорение по итерациям} %%  подпись к рисунку 
+			\label{fig:fig4} 
+		\end{minipage} 
+	\end{center} 
+\end{figure}
+
+Как видно из приведенных данных, присутствует ускорение по времени для задач любой размерности, при этом в целом значение ускорения не линейное, но встречается и сверхлинейное ускорение.
+В данных задачах большое значение имеет что, как быстро алгоритм глобального поиска поставит достаточное для дерева решения количество точек в окрестности глобального минимума, после чего локальный метод быстро находит оптимум, из за этого возникает сверхлинейное ускорение.
+Низкое ускорение в остальных случаях, в большей степени обусловлено тем, что алгоритм оптимизации является синхронным и на некоторых итерациях происходит ожидание завершения локального поиска, также значительным оказалось влияние накладных расходов на передачу точек между процессами.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{credits}
+	\subsubsection{\ackname} A bold run-in heading in small font size at the end of the paper is
+	used for general acknowledgments, for example: This study was funded
+	by X (grant number Y).
+
+	\subsubsection{\discintname}
+	The authors have no competing interests to declare that are
+	relevant to the content of this article.
+\end{credits}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %

### 2025-09-14 2025/MMST_DecisionTree/ParallelDecisionTree.tex
Добален текст про параллельную реализацию

@@
 	\begin{center}
-		\begin{minipage}[h]{0.8\linewidth}
+		\begin{minipage}[h]{0.75\linewidth}
 			\includegraphics[width=1\linewidth]{figure/fig5.png}
@@
 	\begin{center}
-		\begin{minipage}[h]{0.8\linewidth}
+		\begin{minipage}[h]{0.75\linewidth}
 			\includegraphics[width=1\linewidth]{figure/fig4.png}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Одним из способов организации параллельных вычислений на высоко-производительном кластере, который имеет большое количество вычислитель-ных ядер и при этом большое количество потоков на каждом, является синхронное параллельное вычисление значений функции.
+
+В предлагаемом подходе схема распараллеливания соответствует принципу "мастер-работник".
+В мастер-процессе выполняется алгоритм глобального поиска, он накапливает поисковую информацию, на ее основе вычисляет константу Липшица для целевой функции, определяет новые пробные точки и распределяет их между рабочими процессами.
+Рабочие процессы получают точки испытания от главного процесса, вычисляют значение функции в этих точках и отправляют результаты испытаний главному процессу.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 At the initial stage, the master process (let it have the identifier 0) launches $p$ trials in parallel at $p$ different points $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$, where preimages $\{x^1,x^2,\ldots,x^p\}$ are form the interval $[0,1]$.
@@
 %Здесь нужно связующее предложение
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+На каждой итерации мы будем брать несколько точек для вычисления, будем отправлять их на процессы по группам.
+Для этого используется MPI, причем нулевой процесс занимается отправкой точек на вычисление остальным и сбором результатов, но непосредственно в подсчетах не участвует.
+На каждом процессе, кроме нулевого, мы имеем не одну точку, а группу, вычислять которые будем параллельно с использованием центрального процессора.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2025-09-14 2025/MMST_DecisionTree/ParallelDecisionTree.tex
one more part is translated

--- 2025-09-13 2025/MMST_DecisionTree/ParallelDecisionTree.tex
@@
 
-	\item Перенумеруем (по нижними индексами) точки $x^i, 1\leq i\leq k$, а также граничные точки отрезка $[0,1]$ в порядке возрастания координаты  
+	\item Renumber (using subscripts) the points $x^i, 1\leq i\leq k$,  along with the boundary points of the interval $[0, 1]$ in increasing order of their coordinate values
 	\begin{equation} 
-		\label{agp1_sort} 	0=x_0<\ x_1<\ ...\ <x_{k+1}=1. 	
+		\label{agp1_sort} 	0=x_0<\ x_1<\ ...\ <x_{k+1}=1 	
 	\end{equation} 
-	и сопоставить их со значениями $z_i=f(x_i)$.
+	and associate with them the values $z_i=f(x_i)$.
 
-	\item  Вычислить текущую нижнюю оценку $M$ неизвестной константы Гельдера $H$: 
+	\item  Compute the current lower bound $M$ of the unknown Hölder constant $H$: 
 	\begin{equation} 
@@
 	\end{equation} 
-	где $r>1$ — параметр алгоритма.
-Этот параметр управляет надежностью алгоритма: более высокие значения $r$ гарантируют гарантированное нахождение глобального минимума, выбор меньшего значения ускоряет сходимость алгоритма.
+	where $r>1$ is the method parameter.
+This parameter controls the reliability of the algorithm: higher values of $r$ guarantee finding the global minimum, while choosing a smaller value accelerates the convergence of the algorithm.
 
-	\item  Для каждого интервала $(x_{i-1},x_i), 1\leq i\leq k+1,$ вычислить значение $R(i)$, называемое \textit{характеристикой} интервала, по формулам
-
+	\item  For each interval $(x_{i-1},x_i), 1\leq i\leq k+1,$ compute a value $R(i)$, called \textit{the characteristic} of the interval, according to the following formulas:
 	\begin{equation} 
@@
 	\end{equation} 
+	where \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
 
-	где \(\Delta_i=(x_i-x_{i-1})^\frac{1}{N}\).
-
-	\item   Упорядочить характеристики $R\left(i\right),\ 1\leq i \leq k+1,$ в порядке не возрастания
+	\item   Sort the characteristics  $R\left(i\right),\ 1\leq i \leq k+1,$  in non-increasing order
 	\begin{equation} 
@@
 	\end{equation} 
-	и выделить $n$ интервалов с индексами $t_j,\ 1\le\ j\le\ n$  с наибольшими значениями характеристик.
+and select $n$ intervals with indices $t_j, 1 \le j \le n$, that have the largest characteristic values.
 
-	\item Вычислить новые координаты точки испытания $y'^{k+j}=y\left(x^{k+j}\right), \ 1\leq j\leq n$ прообразом которого является $x^{k+j}\in\left(x_{t_j-1},x_{t_j}\right)$, в соответствии с формулой
+	\item Compute the coordinates of the new trial points $y'^{k+j}=y\left(x^{k+j}\right), \ 1\leq j\leq n$, whose preimages are  $x^{k+j}\in\left(x_{t_j-1},x_{t_j}\right)$, according to the following formula:
 	\begin{equation}
@@
 
-	\item Отправить на процесс $j, \ 1\leq j\leq n$ координату $y'^{k+1} ... y'^{k+n}$  и сообщаем процессу, что нужно произвести одиночное испытание.
+	\item Отправить на процесс $j, \ 1\leq j\leq n$ координату $y'^{k+1} ... y'^{k+n}$  и сообщить процессу, что нужно произвести одиночное испытание.
 
-	\item Отправить на процессы $h, \ 1\leq h\leq c$ координаты точек $y^{k+1} ... y^{k+c}$ и сообщаем процессу, что нужно запустить локальный метод.
+	\item Отправить на процессы $h, \ 1\leq h\leq c$ координаты точек $y^{k+1} ... y^{k+c}$ и сообщить процессу, что нужно запустить локальный метод.
 
@@
 
-	\item Строим дерево решений по всей накопленной поисковой информации, получаем аппроксимирующую функцию $\psi(y)$.
+	\item Build a decision tree using all the accumulated search information and obtain the approximating function $\psi(y)$.
 
-	\item Если дерево решений используется впервые, постройте равномерную сетку
+	\item If the decision tree is being used for the first time, construct a uniform grid
 	\begin{displaymath} 
@@
 	\end{displaymath} 
-	где количество узлов сетки является параметром метода.
+	where the number of grid nodes is a parameter of the method.
 
@@
 
-	\item Для всех точек $y'\in V$ найдите точки $y'_q$, ближайшие к $y'$,
+	\item Для всех точек $y'\in V$ найти точки $y'_q$, ближайшие к $y'$,
 
@@
 
-Следующие $p$ испытания выполняются параллельно в точках $x^{k+j},\ 1\leq j\leq p$, вычисляемых по формуле (\ref{agp5_x1}).
-По завершении испытаний результаты этих испытаний сохраняются в базе поисковой информации, а алгоритм переходит к вычислению новых точек испытания.
-Отметим, что, как правило, процесс испытания в прикладных оптимизационных задачах гораздо более затратный по вычислительным ресурсам по сравнению с вычислением координаты точки.
+The following $p$ trials are executed in parallel at the points$x^{k+j},\ 1\leq j\leq p$,  computed using formula  (\ref{agp5_x1}).
+Upon completion of these trials, their results are stored in the search information database, and the algorithm proceeds to compute new trial points.
+Note that, generally, the trial process in applied optimization problems is far more computationally expensive than computing the coordinates of a point.
 
-Алгоритм останавливается в том случае, если выполняется условие \(\Delta_{t_j} < \varepsilon\) хотя бы для одного значения $t_j,\ 1\le\ j\le\ p$, из (\ref{agp4_R_sort} ).
-Этот критерий остановки (наряду с обычным для итерационных методов критерием ограничения числа выполняемых итераций) используется в прикладных задачах оптимизации, в которых точка глобального минимума $y^*$ априори неизвестна.
+The algorithm terminates if the condition \(\Delta_{t_j} < \varepsilon\) is satisfied for at least one value of $t_j,\ 1\le\ j\le\ p$, from (\ref{agp4_R_sort} ).
+This stopping criterion (along with the usual criterion of limiting the number of iterations, common for iterative methods) is used in applied optimization problems where the global minimum point $y^*$ is not known a priori.
 
-При решении тестовых задач, в которых известна точка глобального минимума $y^*$, можно использовать критерий остановки и по попаданию в окрестность глобального минимума.
-В этом случае метод останавливается, если выполняется условие $\left\|y(x_{t_j})-\ y^\ast\right\| < \varepsilon$ для одного $t_j,\ 1\le\ j\le\ p$ из (\ref{agp4_R_sort}).
+When solving test problems where the global minimizer $y^*$ is known, a stopping criterion based on reaching a neighborhood of the global minimum can be used.
+In this case, the method stops if the condition $\left\|y(x_{t_j})-\ y^\ast\right\| < \varepsilon$ is satisfied for one value of $t_j,\ 1\le\ j\le\ p$ from (\ref{agp4_R_sort}).
 
-В качестве окончательной оценки глобального минимума рассматриваемой задачи берется значение 
+The values  
 \begin{equation} 
@@
 \end{equation} 
+are taken as the final estimate of the global minimum for the problem under consideration.
 
-Обоснование такого способа организации вычислений см.
-в \cite{Strongin2000,Barkalov2016}.
-Модификации с учетом наличия ограничений-неравенств, а также информация о производной целевой функции представлены в \cite{Barkalov2002,Gergel1997}.
+See \cite{Strongin2000,Barkalov2016} for the justification of this computational approach.
+Modifications to account for inequality constraints, as well as information about the derivative of the objective function, are presented in \cite{Barkalov2002,Gergel1997}.
 

### 2025-09-14 2025/MMST_DecisionTree/ParallelDecisionTree.tex

@@
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Parallel global optimization algorithm employing decision trees}
 % If the paper title is too long for the running head, you can set
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%Классические алгоритмы локальной оптимизации предназначены для нахождения единственного экстремума (локального минимума или максимума) в окрестности начального приближения.
+В зависимости от используемой информации о целевой функции, эти методы принято классифицировать на:
+
+Classical local optimization algorithms are designed to find a single extremum (local minimum or maximum) in the neighborhood of an initial guess.
+Depending on the information used about the objective function, these methods are typically classified as:
+\begin{itemize}
+	\item \textbf{Zero-order methods}, which use only function values;
+	\item \textbf{First-order methods} (gradient methods), requiring the computation of first derivatives;
+	\item \textbf{Second-order methods} (e.g., Newton's method), which require the computation of the matrix of second derivatives (Hessian).
+\end{itemize}
+
+%Важно подчеркнуть, что применение градиентных методов (первого и, в особенности, второго порядка) для оптимизации функций типа черного ящика сопряжено с фундаментальными трудностями.
+Основная проблема заключается в отсутствии аналитического описания функции, что делает невозможным прямое вычисление производных и приводит к необходимости их численной оценки, что, в свою очередь, может быть вычислительно затратно и вносить дополнительные погрешности \cite{Kelley}.
+В связи с этим, в рамках данного исследования дальнейшее рассмотрение будет ограничено исключительно методами нулевого порядка, не требующими вычисления производных.
+It is important to emphasize that the application of gradient methods (first-order and, especially, second-order) to the optimization of black-box functions is associated with fundamental difficulties.
+The main problem lies in the absence of an analytical description of the function, which makes direct computation of derivatives impossible and leads to the need for their numerical estimation.
+This, in turn, can be computationally expensive and introduce additional errors \cite{Kelley}.
+In light of this, the remainder of this study will be limited exclusively to zero-order methods, which do not require the computation of derivatives.
+
+%Ключевой вопрос при гибридизации глобального и локального поиска заключается в определении оптимального момента для запуска локального метода.
+Для решения этой задачи предлагается использовать адаптивный алгоритм, основанный на машином обучении, а именно на построении дерева решений.
+%На первоначальном этапе глобального поиска информация о поведении целевой функции накапливается в виде набора вычисленных значений в точках испытаний.
+Эти данные используются для обучения модели дерева решений.
+Обученная модель строит кусочно-постоянную аппроксимацию целевой функции, которая позволяет прогнозировать её поведение в ещё не исследованных областях.
+%Механизм принятия решения о запуске локального поиска заключается в следующем: для точки, подозрительной на наличие в её окрестности локального минимума, с помощью аппроксимирующей модели анализируются значения в соседних точках.
+На основе этого анализа делается вывод о том, попадает ли следующая точка испытания глобального алгоритма в область притяжения данного локального минимума.
+Это позволяет своевременно запустить локальную оптимизацию для эффективного уточнения решения в данной области.
+
+A key question when hybridizing global and local search is determining the optimal moment to initiate the local method.
+To address this, we propose using an adaptive algorithm based on machine learning, specifically on building a decision tree.
+
+In the initial phase of the global search, information about the behavior of the objective function is accumulated in the form of a set of computed values at trial points.
+This data is used to train a decision tree model.
+The trained model constructs a piecewise-constant approximation of the objective function, which allows predicting its behavior in unexplored regions.
+
+The mechanism for deciding when to initiate local search is as follows: for a point suspected of having a local minimum in its vicinity, the approximating model is used to analyze values in neighboring points.
+Based on this analysis, a conclusion is drawn as to whether the next trial point of the global algorithm falls within the basin of attraction of this local minimum.
+This allows timely initiation of local optimization for effective refinement of the solution in that region.
+
+The content of this paper is structured in the following manner. 
+Section \ref{SecA} addresses the methodological basis of the research.
+It outlines the main idea of the global search algorithm used and details the computational rules of its operation.
+Furthermore, the section includes a description of the employed local optimization method (zero-order) and justifies the technique for constructing a function approximation using decision trees.
+Section \ref{SecGSA} presents the key contribution of this work: a novel hybrid algorithm integrating global search with local refinement, managed via predictions from a decision tree model.
+Section \ref{SecR} contains the results of computational experiments conducted on a parallel computing system, their analysis and interpretation, and the formulation of the main concluding remarks.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Problem statement and solution strategies}\label{SecA}
@@
 Consequently, rather than addressing the original problem of minimizing the function $\varphi(y)$ within the domain $D$, we can consider minimizing the one-dimensional function $f(x)=\varphi(y(x))$, subject to the Hölder condition for $x \in [0, 1]$.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+This approach is discussed in the works \cite{Sergeyev2022, Usova2024}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 	\begin{center}
-		\begin{minipage}[h]{0.7\linewidth}
+		\begin{minipage}[h]{0.75\linewidth}
 			\includegraphics[width=1\linewidth]{figure/fig5.png}
@@
 	\begin{center}
-		\begin{minipage}[h]{0.7\linewidth}
+		\begin{minipage}[h]{0.75\linewidth}
 			\includegraphics[width=1\linewidth]{figure/fig4.png}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The exploratory search is conducted as follows:
+\begin{itemize}[label=$\bullet$] 
+	\item The step size is determined (it is different for each coordinate direction and can change during the search);
+	\item A search step is considered successful if the objective function value at the new point does not exceed the value at the initial point; 
+	\item Otherwise, it is necessary to return and take a step in the opposite direction; 
+	\item After iterating through all $N$ coordinates, the exploratory search ends; the resulting point is called the base point.
+\end{itemize}
+The pattern search consists of taking one step from the resulting base point along the line connecting it to the previous base point.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{figure} 
 	\begin{center} 
-		\begin{minipage}[h]{0.8\linewidth} 
+		\begin{minipage}[h]{1.0\linewidth} 
 			\includegraphics[width=1\linewidth]{figure/fig1.pdf} 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+One way to organize parallel computations on a supercomputer, which has a large number of computational cores and a high thread count per core, is to use synchronous parallel function evaluation.
+
+In the proposed approach, the parallelization scheme follows a ``master-worker'' principle.
+The master process executes the global search algorithm, accumulates search information, calculates the Lipschitz constant for the objective function based on this information, determines new trial points, and distributes them among the worker processes.
+The worker processes receive the trial points from the master process, evaluate the function at those points, and send the results back to the master process.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 At the initial stage, the master process (let it have the identifier 0) launches $p$ trials in parallel at $p$ different points $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$, where preimages $\{x^1,x^2,\ldots,x^p\}$ are form the interval $[0,1]$.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In each iteration, the algorithm selects several computation points and distributes them to processes in groups.
+MPI manages this, with process zero handling point distribution and result collection, but remaining computationally idle.
+Each remaining process receives a group of points, which are computed in parallel using the CPU.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{enumerate}
 
-	\item Извлекаем из множество $U$ (координат точек испытания) $c = max(p, |U|)$ координат точек испытания $y^{k+1}=y\left(x^{k+1}\right) ...
+	\item Извлекаем из множества $U$ (координат точек испытания) $c = max(p, |U|)$ координат точек испытания $y^{k+1}=y\left(x^{k+1}\right) ...
 y^{k+c}=y\left(x^{k+c}\right)$.
@@
 See \cite{Strongin2000,Barkalov2016} for the justification of this computational approach.
-Modifications to account for inequality constraints, as well as information about the derivative of the objective function, are presented in \cite{Barkalov2002,Gergel1997}.
+Modifications to account for inequality constraints, as well as information about the derivative of the objective function, are presented in \cite{Barkalov2002, Gergel1997, Barkalov2023, Gegrel2021}.
 
@@
 
-Критерием остановки алгоритмов было попадание очередных точек испытаний в $\varepsilon$-окрестность истинного глобального минимума.
-В таблицах \ref{tab:1} \ref{tab:2} сравниваются усредненные числа итераций, выполняемых алгоритмами при решении задач, и значения ускорения времени относительно последовательного прогона.
-Распараллеливание выполнено по технологии MPI, эксперимент выполнен на 8, 16, 32 и 64 процессах.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Algorithms terminated when current trial point was within the $\varepsilon$-neighborhood of the true global minimum.
+Tables \ref{tab:1} and \ref{tab:2} show the average solution time and speedup relative to a sequential run, achieved by the algorithms when solving optimization problems.
+Figure \ref{fig:fig4} shows the speedup in terms of the number of iterations.
+Parallelization was implemented using MPI, and experiments were conducted on 1, 8, 16, 32, and 64 processes.
 
 \begin{table}[!ht]
-	\caption{Ускорение, выполненных разными алгоритмами}
+	\caption{The average solution time}
+	\label{tab:1}
+	\centering	
+	\begin{tabular}{|l|l|l|l|l|l|}
+		\hline
+		2 & 1 mpi & 8 mpi & 16 mpi & 32 mpi & 64 mpi  \\ \hline
+		hard & 12,71 & 1,98 & 1,71 & 0,87 & 1,65  \\ \hline
+		simple & 6,35 & 0,57 & 1,27 & 0,95 & 0,21  \\ \hline
+		3 & ~ & ~ & ~ & ~ &  \\ \hline
+		hard & 14,19 & 5,59 & 2,10 & 0,75 & 0,60  \\ \hline
+		simple & 5,92 & 1,63 & 1,17 & 0,38 & 0,41  \\ \hline
+		4 & ~ & ~ & ~ & ~ &   \\ \hline
+		hard & 66,29 & 34,93 & 8,53 & 4,95 & 2,93  \\ \hline
+		simple & 31,88 & 9,44 & 4,71 & 4,67 & 5,51  \\ \hline
+		5 & ~ & ~ & ~ & ~ &  \\ \hline
+		hard & 324,04 & 52,42 & 17,55 & 11,48 & 10,60 \\ \hline
+		simple & 53,77 & 6,62 & 3,38 & 1,99 & 1,70  \\ \hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[!ht]
+	\caption{Time speedup}
 	\label{tab:2}
@@
 
-Как видно из таблицы \ref{tab:2}, ускорение итерации достаточно велико для задач любой размерности.
+%\begin{table}[!ht]
+%	\caption{Ускорениепо итерациям}
+%	\label{tab:3}
+%	\centering	
+%	\begin{tabular}{|l|l|l|l|l|}
+	%		\hline
+	%		2 & 8 mpi & 16 mpi & 32 mpi & 64 mpi  \\ \hline
+	%		hard & 37,61 & 49,11 & 50,04 & 34,58  \\ \hline
+	%		simple & 52,06 & 19,72 & 63,02 & 307,59  \\ \hline
+	%		3 & ~ & ~ & ~ &   \\ \hline
+	%		hard & 15,03 & 92,04 & 47,28 & 89,74  \\ \hline
+	%		simple & 10,25 & 44,70 & 44,29 & 76,87  \\ \hline
+	%		4 & ~ & ~ & ~ &   \\ \hline
+	%		hard & 38,14 & 106,64 & 58,40 & 117,56  \\ \hline
+	%		simple & 30,75 & 79,84 & 29,41 & 40,43  \\ \hline
+	%		5 & ~ & ~ & ~ &   \\ \hline
+	%		hard & 162,66 & 269,09 & 201,95 & 206,94  \\ \hline
+	%		simple & 46,13 & 176,70 & 98,42 & 191,00  \\ \hline
+	%	\end{tabular}
+%\end{table}
+
+\begin{figure} 
+	\begin{center} 
+		\begin{minipage}[h]{0.9\linewidth} 
+			\includegraphics[width=1\linewidth]{figure/fig6.pdf} 
+			\caption{Iteration speedup} %%  подпись к рисунку 
+			\label{fig:fig4} 
+		\end{minipage} 
+	\end{center} 
+\end{figure}
+
+%Как видно из приведенных данных, присутствует ускорение по времени для задач любой размерности, при этом в целом значение ускорения не линейное, но встречается и сверхлинейное ускорение.
+В данных задачах большое значение имеет что, как быстро алгоритм глобального поиска поставит достаточное для дерева решения количество точек в окрестности глобального минимума, после чего локальный метод быстро находит оптимум, из за этого возникает сверхлинейное ускорение.
+Низкое ускорение в остальных случаях, в большей степени обусловлено тем, что алгоритм оптимизации является синхронным и на некоторых итерациях происходит ожидание завершения локального поиска, также значительным оказалось влияние накладных расходов на передачу точек между процессами.
+
+As can be seen from the presented data, there is a time speedup for problems of all dimensions.
+While the overall speedup is not linear, superlinear speedup does occur.
+In these problems, it is crucial how quickly the global search algorithm places enough points in the neighborhood of the global minimum for the decision tree; after this, the local method quickly finds the optimum, resulting in superlinear speedup.
+Lower speedup in other cases is primarily due to the synchronous nature of the optimization algorithm, leading to waiting for the completion of local searches in some iterations.
+The overhead of transferring points between processes also had a significant impact.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-На данный момент в результате проделанной работы удалось совместить алгоритм глобального поиска с методом локальной оптимизации.
-В отличие от известных многозаходных схем решение о запуске локального метода принимается с помощью дерева решений.
+%На данный момент в результате проделанной работы удалось совместить алгоритм глобального поиска с эффективным локальным уточнением.
+В отличие от известных мультистартовых схем решение о запуске локального метода принимается с помощью дерева решений.
 Использование такой комбинации методов позволяет значительно ускорить работу алгоритма.
-
-Параллельная версия алгоритма сохраняет свойства своего последовательного прототипа, что было подтверждено численными экспериментами по решению серии из нескольких сотен задач различной размерности.
+%Параллельная версия алгоритма сохраняет свойства своего последовательного прототипа, что было подтверждено численными экспериментами по решению серии из нескольких сотен задач различной размерности.
 Предлагаемая схема позволяет использовать преимущества как распараллеливания, так и быстрого поиска локальных экстремумов.
-
-Помимо использования деревьев решений для выявления областей притяжения локальных экстремумов многоэкстремальных функций, мы также планируем использовать методы машинного обучения для разделения переменных решаемой задачи.
+%Помимо использования деревьев решений для выявления областей притяжения локальных экстремумов многоэкстремальных функций, мы также планируем использовать методы машинного обучения для разделения переменных решаемой задачи.
 Во многих прикладных задачах оптимизации зависимость целевой функции от некоторых параметров бывает либо линейной, либо одномодальной.
 Выделение таких переменных в специальную группу и решение задачи с помощью параллельной схемы рекурсивной оптимизации \cite{Barkalov2020_1} позволяет сократить время решения задачи на порядки по сравнению с использованием глобального поиска сразу по всем переменным.
+
+Currently, the work done has made it possible to combine a global search algorithm with effective local refinement.
+Unlike known multi-start schemes, the decision to launch the local method is made using a decision tree.
+The use of such a combination of methods significantly speeds up the algorithm.
+
+The parallel version of the algorithm retains the properties of its sequential prototype, which was confirmed by numerical experiments on solving a series of several hundred problems of various dimensions.
+The proposed scheme allows us to exploit the advantages of both parallelization and rapid local extrema search.
+
+In addition to using decision trees to identify the basins of attraction of local extrema of multimodal functions, we also plan to use machine learning methods to separate the variables of the problem being solved.
+In many applied optimization problems, the dependence of the objective function on some parameters is either linear or unimodal.
+Identifying such variables into a special group and solving the problem using a parallel recursive optimization scheme \cite{Barkalov2020_1} can reduce the solution time by orders of magnitude compared to using global search across all variables at once.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{credits}
+	\subsubsection{\ackname} This study was supported by the Ministry of Science and Higher Education of the Russian Federation, project no. FSWR-2023-0034.
+
+	\subsubsection{\discintname}
+	The authors have no competing interests to declare that are	relevant to the content of this article.
+\end{credits}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2025-09-15 2025/MMST_DecisionTree/ParallelDecisionTree.tex
Поправлен перевод

@@
 
-\usepackage[utf8]{inputenc} % кодировка
-\usepackage[english, russian]{babel} % Русские и английские переносы
+%\usepackage[utf8]{inputenc} % кодировка
+\usepackage[english]{babel} % Русские и английские переносы
 \usepackage{cite}              % для корректного оформления литературы
@@
 	\begin{center}
-		\begin{minipage}[h]{0.75\linewidth}
+		\begin{minipage}[h]{0.7\linewidth}
 			\includegraphics[width=1\linewidth]{figure/fig5.png}
@@
 	\begin{center}
-		\begin{minipage}[h]{0.75\linewidth}
+		\begin{minipage}[h]{0.7\linewidth}
 			\includegraphics[width=1\linewidth]{figure/fig4.png}
@@
 	\begin{center} 
-		\begin{minipage}[h]{1.0\linewidth} 
+		\begin{minipage}[!h]{1.0\linewidth} 
 			\includegraphics[width=1\linewidth]{figure/fig1.pdf} 
@@
 
-	\item Извлекаем из множества $U$ (координат точек испытания) $c = max(p, |U|)$ координат точек испытания $y^{k+1}=y\left(x^{k+1}\right) ...
-y^{k+c}=y\left(x^{k+c}\right)$.
-Если $c < p$ то сначала вычисляем $n = p - c$ точек согласно шагам 2 - 7, иначе сразу переходим к шагу 8.
+	\item Extract $c = \max(p, |U|)$ trial point coordinates $y^{k+1}=y\left(x^{k+1}\right) ...
+y^{k+c}=y\left(x^{k+c}\right)$ from the set $U$ (which contains trial point coordinates).
+If $c < p$, then first perform trials at $n = p - c$ points according to Steps~2--7; otherwise proceed directly to Step~8.
 
@@
 
-	\item Отправить на процесс $j, \ 1\leq j\leq n$ координату $y'^{k+1} ... y'^{k+n}$  и сообщить процессу, что нужно произвести одиночное испытание.
+	\item Send to process $j$, $1 \leq j \leq n$, the coordinates $y'^{k+1}, \dots, y'^{k+n}$ and notify the process that it needs to perform a single trial.
 
-	\item Отправить на процессы $h, \ 1\leq h\leq c$ координаты точек $y^{k+1} ... y^{k+c}$ и сообщить процессу, что нужно запустить локальный метод.
+	\item Send to processes $h$, $1 \leq h \leq c$, the coordinates of points $y^{k+1}, \dots, y^{k+c}$ and notify the process that it needs to run a local method.
 
-	\item Получить со всех $p$ процессов значения испытаний и добавить их в поисковую информацию.
+	\item Receive trial values from all $p$ processes and add them to the search information.
 
-	\item Добавить точки испытаний $y'^{k+1}=y\left(x^{k+1}\right) ... y'^{k+n}=y\left(x^{k+т}\right)$ к множеству $V$.
+	\item Add the trial points $y'^{k+1}=y\left(x^{k+1}\right) , \dots, y'^{k+n}=y\left(x^{k+n}\right)$  to the set $V$.
 
-	\item Если $k\ <\ 100 N$, перейти к шагу 1.
+	\item If $k < 100N$, go to Step~1.
 
@@
 
-	\item Вычислить кусочно постоянную аппроксимацию функции $\psi(y)$, по дереву решений : $Z' = \{ z'=  \psi(y'), y' \in Y'\}$
+	\item Compute the piecewise constant approximation of the function $\psi(y)$ using the decision tree: $Z' = \{ z' = \psi(y'), y' \in Y' \}$.
 
-	\item Для всех точек $y'\in V$ найти точки $y'_q$, ближайшие к $y'$,
+	\item For all points $y' \in V$, find the points $y'_q$ closest to $y'$.
 
-	Если ни одна точка $y'_q$ не имеет меньшего значения, чем $y'_q$, поместить $y'$ в множество $U$
+	If no point $y'_q$ has a value smaller than $y'$, add $y'$ to the set $U$.
 
-	Очистить множество $V$.
+	Clear the set $V$.
 
-	\item Проверка критериев остановки
+	\item Check the stopping criteria.
 
@@
 Parallelization was implemented using MPI, and experiments were conducted on 1, 8, 16, 32, and 64 processes.
+
+\begin{figure} 
+	\begin{center} 
+		\begin{minipage}[h]{0.9\linewidth} 
+			\includegraphics[width=1\linewidth]{figure/fig6.pdf} 
+			\caption{Iteration speedup} %%  подпись к рисунку 
+			\label{fig:fig4} 
+		\end{minipage} 
+	\end{center} 
+\end{figure}
 
@@
 
-\begin{figure} 
-	\begin{center} 
-		\begin{minipage}[h]{0.9\linewidth} 
-			\includegraphics[width=1\linewidth]{figure/fig6.pdf} 
-			\caption{Iteration speedup} %%  подпись к рисунку 
-			\label{fig:fig4} 
-		\end{minipage} 
-	\end{center} 
-\end{figure}
-
 %Как видно из приведенных данных, присутствует ускорение по времени для задач любой размерности, при этом в целом значение ускорения не линейное, но встречается и сверхлинейное ускорение.

### 2025-09-15 2025/MMST_DecisionTree/ParallelDecisionTree.tex
Поправлен текст

@@
 
-	\item Send to process $j$, $1 \leq j \leq n$, the coordinates $y'^{k+1}, \dots, y'^{k+n}$ and notify the process that it needs to perform a single trial.
+	\item Send to process $j$, $1 \leq j \leq n$, the coordinates of points $y'^{k+1}, \dots, y'^{k+n}$ and notify the process that it needs to perform a single trial.
 

### 2025-09-16 2025/MMST_DecisionTree/ParallelDecisionTree.tex
Поправлено оформление

@@
 Ilya Lebedev\orcidID{0000-0002-8736-0652} \and
-Dmitry Silenko}
+Dmitry Silenko\orcidID{0000-0002-2578-9699} }
 %
@@
 
-\begin{table}[!ht]
+\begin{table}
 	\caption{The average solution time}
 	\label{tab:1}
-	\centering	
-	\begin{tabular}{|l|l|l|l|l|l|}
-		\hline
-		2 & 1 mpi & 8 mpi & 16 mpi & 32 mpi & 64 mpi  \\ \hline
-		hard & 12,71 & 1,98 & 1,71 & 0,87 & 1,65  \\ \hline
-		simple & 6,35 & 0,57 & 1,27 & 0,95 & 0,21  \\ \hline
-		3 & ~ & ~ & ~ & ~ &  \\ \hline
-		hard & 14,19 & 5,59 & 2,10 & 0,75 & 0,60  \\ \hline
-		simple & 5,92 & 1,63 & 1,17 & 0,38 & 0,41  \\ \hline
-		4 & ~ & ~ & ~ & ~ &   \\ \hline
-		hard & 66,29 & 34,93 & 8,53 & 4,95 & 2,93  \\ \hline
-		simple & 31,88 & 9,44 & 4,71 & 4,67 & 5,51  \\ \hline
-		5 & ~ & ~ & ~ & ~ &  \\ \hline
-		hard & 324,04 & 52,42 & 17,55 & 11,48 & 10,60 \\ \hline
-		simple & 53,77 & 6,62 & 3,38 & 1,99 & 1,70  \\ \hline
+	\center
+	\begin{tabular}{cccccccccccc}
+		\hline\noalign{\smallskip}
+		$P$ & \multicolumn{2}{c}{ $N=2$ } & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
+		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12} \noalign{\smallskip}
+		& \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
+		\noalign{\smallskip} \hline \noalign{\smallskip}
+		1   & 6,35 & 12,71 & & 5,92 & 14,19 & & 31,88 & 66,29  & & 53,77 & 324,04 \\
+		8   & 0,57 & 1,98  & & 1,63 & 5,59  & & 9,44  & 34,93  & & 6,62 & 52,42 \\
+		16  & 1,27 & 1,71  & & 1,17 & 2,10  & & 4,71  & 8,53   & & 3,38 & 17,55 \\
+		32  & 0,95 & 0,87  & & 0,38 & 0,75  & & 4,67  & 4,95   & & 1,99 & 11,48 \\
+		64  & 0,21 & 1,65  & & 0,41 & 0,60  & & 5,51  & 2,93   & & 1,70 & 10,60 \\
+		\noalign{\smallskip}\hline
 	\end{tabular}
@@
 
-\begin{table}[!ht]
+\begin{table}
 	\caption{Time speedup}
 	\label{tab:2}
-	\centering	
-	\begin{tabular}{|l|l|l|l|l|}
-		\hline
-		2 & 8 mpi & 16 mpi & 32 mpi & 64 mpi  \\ \hline
-		hard & 6,42 & 7,43 & 16,45 & 7,70  \\ \hline
-		simple & 11,19 & 5,018 & 13,59 & 30,42  \\ \hline
-		3 & ~ & ~ & ~ &   \\ \hline
-		hard & 4,01 & 6,76 & 22,41 & 23,48  \\ \hline
-		simple & 3,98 & 5,07 & 14,97 & 14,49  \\ \hline
-		4 & ~ & ~ & ~ &   \\ \hline
-		hard & 3,52 & 7,77 & 15,64 & 22,61  \\ \hline
-		simple & 4,20 & 6,77 & 7,27 & 5,79  \\ \hline
-		5 & ~ & ~ & ~ &   \\ \hline
-		hard & 10,28 & 18,46 & 38,76 & 30,57  \\ \hline
-		simple & 10,80 & 15,93 & 31,96 & 31,65  \\ \hline
+	\center
+	\begin{tabular}{cccccccccccc}
+		\hline\noalign{\smallskip}
+		$P$ & \multicolumn{2}{c}{ $N=2$ } & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
+		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12} \noalign{\smallskip}
+		& \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
+		\noalign{\smallskip} \hline \noalign{\smallskip}
+		8  & 11,19 & 6,42  & & 3,98  & 4,01  & & 4,20 & 3,52  & & 10,80 & 10,28 \\
+		16 & 5,02  & 7,43  & & 5,07  & 6,76  & & 6,77 & 7,77  & & 15,93 & 18,46 \\
+		32 & 13,59 & 16,45 & & 14,97 & 22,41 & & 7,27 & 15,64 & & 31,96 & 38,76 \\
+		64 & 30,42 & 7,70  & & 14,49 & 23,48 & & 5,79 & 22,61 & & 31,65 & 30,57 \\
+		\noalign{\smallskip}\hline
 	\end{tabular}

### 2025-09-17 2025/MMST_DecisionTree/ParallelDecisionTree.tex
minor corrections in formulas

--- 2025-09-15 2025/MMST_DecisionTree/ParallelDecisionTree.tex
@@
 
-	\item Extract $c = \max(p, |U|)$ trial point coordinates $y^{k+1}=y\left(x^{k+1}\right) ...
-y^{k+c}=y\left(x^{k+c}\right)$ from the set $U$ (which contains trial point coordinates).
+	\item Extract $c = \max(p, |U|)$ trial point coordinates $y^{k+1}=y\left(x^{k+1}\right),...,y^{k+c}=y\left(x^{k+c}\right)$ from the set $U$ (which contains trial point coordinates).
 If $c < p$, then first perform trials at $n = p - c$ points according to Steps~2--7; otherwise proceed directly to Step~8.
@@
 		\label{agp5_x1}
-		x^{k+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}\left(z_t-z_{t-1}\right)\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{M}\right]^N.
+		x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}\left(z_{t_j}-z_{t_j-1}\right)\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{M}\right]^N.
 	\end{equation}
@@
 
-	\item Receive trial values from all $p$ processes and add them to the search information.
+	\item Receive trial results from all $p$ processes and add them to the search information.
 
@@
 	\item For all points $y' \in V$, find the points $y'_q$ closest to $y'$.
-
 		If no point $y'_q$ has a value smaller than $y'$, add $y'$ to the set $U$.
-
 		Clear the set $V$.

### 2025-09-17 2025/MMST_DecisionTree/ParallelDecisionTree.tex

@@
 The Global Search Algorithm (GSA) is based on this idea \cite{Strongin2000}.
-In this work, we attempted to combine the GSA and a local optimization method \cite{Nocedal, Kelley} (the Hooke-Jeeves pattern search method \cite{HookJeeves}) to reduce the number of trials performed.
-The decision to initiate the local method will be made using a decision tree.
+In this work, we attempted to combine the GSA and a local optimization method (the Hooke-Jeeves pattern search method \cite{Nocedal, Kelley}) to reduce the number of trials performed.
+The decision to run the local method will be made using a decision tree.
 
@@
 \begin{itemize}
-	\item \textbf{Zero-order methods}, which use only function values;
-	\item \textbf{First-order methods} (gradient methods), requiring the computation of first derivatives;
-	\item \textbf{Second-order methods} (e.g., Newton's method), which require the computation of the matrix of second derivatives (Hessian).
+	\item Zero-order methods, which use only function values;
+	\item First-order methods (gradient methods), requiring the computation of first derivatives;
+	\item Second-order methods (e.g., Newton's method), which require the computation of the matrix of second derivatives (Hessian).
 \end{itemize}
@@
 
-Algorithms terminated when current trial point was within the $\varepsilon$-neighborhood of the true global minimum.
-Tables \ref{tab:1} and \ref{tab:2} show the average solution time and speedup relative to a sequential run, achieved by the algorithms when solving optimization problems.
+Algorithms terminated when current trial point was within the $\varepsilon$-vicinity of the true global minimizer.
+Tables \ref{tab:1} and \ref{tab:2} show the average time and speedup relative to a sequential run, achieved by the algorithms when solving optimization problems.
 Figure \ref{fig:fig4} shows the speedup in terms of the number of iterations.
-Parallelization was implemented using MPI, and experiments were conducted on 1, 8, 16, 32, and 64 processes.
+Parallelization was implemented using MPI, and experiments were conducted on $P = 1, 8, 16, 32, 64$ processes.
 
@@
 \begin{table}
-	\caption{The average solution time}
+	\caption{Average time to solve the problem}
 	\label{tab:1}
@@
 		\noalign{\smallskip} \hline \noalign{\smallskip}
-		1   & 6,35 & 12,71 & & 5,92 & 14,19 & & 31,88 & 66,29  & & 53,77 & 324,04 \\
-		8   & 0,57 & 1,98  & & 1,63 & 5,59  & & 9,44  & 34,93  & & 6,62 & 52,42 \\
-		16  & 1,27 & 1,71  & & 1,17 & 2,10  & & 4,71  & 8,53   & & 3,38 & 17,55 \\
-		32  & 0,95 & 0,87  & & 0,38 & 0,75  & & 4,67  & 4,95   & & 1,99 & 11,48 \\
-		64  & 0,21 & 1,65  & & 0,41 & 0,60  & & 5,51  & 2,93   & & 1,70 & 10,60 \\
+		1   & 6.35 & 12.71 & & 5.92 & 14.19 & & 31.88 & 66.29  & & 53.77 & 324.04 \\
+		8   & 0.57 & 1.98  & & 1.63 & 5.59  & & 9.44  & 34.93  & & 6.62 & 52.42 \\
+		16  & 1.27 & 1.71  & & 1.17 & 2.10  & & 4.71  & 8.53   & & 3.38 & 17.55 \\
+		32  & 0.95 & 0.87  & & 0.38 & 0.75  & & 4.67  & 4.95   & & 1.99 & 11.48 \\
+		64  & 0.21 & 1.65  & & 0.41 & 0.60  & & 5.51  & 2.93   & & 1.70 & 10.60 \\
 		\noalign{\smallskip}\hline
@@
 		\noalign{\smallskip} \hline \noalign{\smallskip}
-		8  & 11,19 & 6,42  & & 3,98  & 4,01  & & 4,20 & 3,52  & & 10,80 & 10,28 \\
-		16 & 5,02  & 7,43  & & 5,07  & 6,76  & & 6,77 & 7,77  & & 15,93 & 18,46 \\
-		32 & 13,59 & 16,45 & & 14,97 & 22,41 & & 7,27 & 15,64 & & 31,96 & 38,76 \\
-		64 & 30,42 & 7,70  & & 14,49 & 23,48 & & 5,79 & 22,61 & & 31,65 & 30,57 \\
+		8  & 11.19 & 6.42  & & 3.98  & 4.01  & & 4.20 & 3.52  & & 10.80 & 10.28 \\
+		16 & 5.02  & 7.43  & & 5.07  & 6.76  & & 6.77 & 7.77  & & 15.93 & 18.46 \\
+		32 & 13.59 & 16.45 & & 14.97 & 22.41 & & 7.27 & 15.64 & & 31.96 & 38.76 \\
+		64 & 30.42 & 7.70  & & 14.49 & 23.48 & & 5.79 & 22.61 & & 31.65 & 30.57 \\
 		\noalign{\smallskip}\hline

### 2025-09-19 2025/RuSCDays/RSD-2025.tex
delete colontitules and add all emails

@@
 %
-\titlerunning{A Parallel Algorithm for Tuning Hyperparameters}
+\titlerunning{ }%A Parallel Algorithm for Tuning Hyperparameters}
 % If the paper title is too long for the running head, you can set
@@
 \author{Marina Usova \letter \orcidID{0000-0002-0722-6884} \and
-Ilya Lebedev\orcidID{0000-0002-8736-0652}  \and
-Anton Shtanyuk\orcidID{0000-0003-1809-7173} \and
-Evgeny Kozinov \orcidID{0000-0001-6776-0096} \and
-Konstantin Barkalov \orcidID{0000-0001-5273-2471} 
+Ilya Lebedev \letter \orcidID{0000-0002-8736-0652}  \and
+Anton Shtanyuk \letter \orcidID{0000-0003-1809-7173} \and
+Evgeny Kozinov \letter \orcidID{0000-0001-6776-0096} \and
+Konstantin Barkalov \letter \orcidID{0000-0001-5273-2471} 
 }
 %
-\authorrunning{M. Usova et al.}
+\authorrunning{ }%M. Usova et al.}
 % First names are abbreviated in the running head.
@@
 \institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
-\email{usova@itmm.unn.ru}}
+\email{usova@itmm.unn.ru},
+\email{ilya.lebedev@itmm.unn.ru},
+\email{anton.shtanyuk@itmm.unn.ru},
+\email{evgeny.kozinov@itmm.unn.ru},
+\email{konstantin.barkalov@itmm.unn.ru}}
 %

### 2025-09-19 2025/RSCD_MCO/RSD-2025.tex
fix RSCD editor comments

@@
 %
-\title{Parallel algorithm for solving multicriterial optimization problems using elements of machine learning}
+\title{Parallel Algorithm for Solving Multicriterial Optimization Problems Using Elements of Machine Learning}
 %
-\titlerunning{Parallel algorithm for solving multicriterial optimization problems}
+\titlerunning{  %Parallel Algorithm for Solving Multicriterial Optimization Problems
+}
 % If the paper title is too long for the running head, you can set
@@
 %
-\authorrunning{S. Konnov et al.}
+\authorrunning{
+%S. Konnov et al.
+}
 % First names are abbreviated in the running head.
@@
 
-\section{Problem statement}
+\section{Problem Statement}
 \label{sec2}
@@
 
-\section{Computational scheme of the developed algorithm}
+\section{Computational Scheme of the Developed Algorithm}
 \label{sec3}
@@
 
-\subsection{Block of the global search algorithm}
+\subsection{Block of the Global Search Algorithm}
 \label{subsec31}
@@
 
-\subsection{Block for accumulating and reusing the information about current state of the search}
+\subsection{Block for Accumulating and Reusing the Information About Current State of the Search}
 \label{subsec32}
@@
 
-\subsection{Block for building the Pareto set estimation}
+\subsection{Block for Building the Pareto Set Estimation}
 \label{subsec33}
@@
 
-\subsection{Block of  building the machine learning for estimation of the Pareto set}
+\subsection{Block of Building the Machine Learning for Estimation of the Pareto Set}
 \label{subsec34}
@@
 
-\subsection{Block for the choice of trial points and implementation of computational experiments}
+\subsection{Block for the Choice of Trial Points and Implementation of Computational Experiments}
 \label{subsec35}
@@
 
-\section{Results of computational experiments}
+\section{Results of Computational Experiments}
 \label{sec4}

### 2025-10-08 2025/MMST_DecisionTree/ParallelDecisionTree.tex
first minor revision

@@
 %
-\author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
-Ilya Lebedev\orcidID{0000-0002-8736-0652} \and
-Dmitry Silenko\orcidID{0000-0002-2578-9699} }
+\author{Konstantin Barkalov$^{1,2}$\orcidID{0000-0001-5273-2471} \and
+Ilya Lebedev$^{1,2}$\orcidID{0000-0002-8736-0652} \and
+Dmitry Silenko$^{1,2}$\orcidID{0000-0002-2578-9699} }
 %
@@
 %
-\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia}
+\institute{$^1$Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia\\$^2$Research and Education Mathematics Center ``Mathematics for Future Technologies'', Nizhni Novgorod, Russia}
 %
@@
 This paper investigates parallel algorithms for solving multidimensional global optimization problems These problems commonly occur when optimizing the parameters of a mathematical model to achieve the best possible agreement between simulation results and experimental observations.
-Numerous algorithms have been developed to address this class of problems, varying from metaheuristic techniques based on the idea of random search \cite{Ferreiro2013,Garcia2014,Langdon2011} to deterministic algorithms designed to guarantee convergence to a global minimum \cite{Evtushenko2009,He2008,Paulavicius2011}.
+Numerous algorithms have been developed to address this class of problems, varying from metaheuristic and evolutionary techniques \cite{Ferreiro2013,Garcia2014,Langdon2011} to deterministic algorithms designed to guarantee convergence to a global minimizer \cite{Evtushenko2009,He2008,Paulavicius2011}.
 
@@
 This may be achieved by the purposeful selection of alternatives during the course of optimization, rejecting unpromising search sub-regions and exploring only those where a solution can be found.
-The Global Search Algorithm (GSA) is based on this idea \cite{Strongin2000}.
-In this work, we attempted to combine the GSA and a local optimization method (the Hooke-Jeeves pattern search method \cite{Nocedal, Kelley}) to reduce the number of trials performed.
+%именно на такой идее основан один из эффективных детерминированных методов глобальной оптимизации - Алгоритм Глобального поиска
+This idea forms the foundation of an efficient deterministic global optimization method: the Global Search Algorithm (GSA) \cite{Strongin2000}.
+In this work, we attempted to combine the GSA and a local optimization method (the Hooke-Jeeves method \cite{Nocedal, Kelley}) to decrease the number of trials carried out.
 The decision to run the local method will be made using a decision tree.
+
+% Следует отметить, что и другие подходы к решению задач глобальной оптмизации используют схожие идеи, основанные на использования локального уточнения лучшего найденного решения для ускорения сходимости к точке глобального оптимума.
+Сказанное справедливо как для эволюционных, так и для детерминированных алгоритмов глобальной оптмизации.
+
+It should be noted that other approaches to solving global optimization problems use similar ideas based on employing local refinement of the best-found solution to accelerate convergence to the global optimizer.
+This holds true for both evolutionary \cite{Kuzenkov2024} and deterministic \cite{Grishagin2025} global optimization algorithms.
 
@@
 
+%Первая версия заключения
+
 %На данный момент в результате проделанной работы удалось совместить алгоритм глобального поиска с эффективным локальным уточнением.
@@
 
+%Currently, the work done has made it possible to combine a global search algorithm with effective local refinement.
+Unlike known multi-start schemes, the decision to launch the local method is made using a decision tree.
+The use of such a combination of methods significantly speeds up the algorithm.
+%The parallel version of the algorithm retains the properties of its sequential prototype, which was confirmed by numerical experiments on solving a series of several hundred problems of various dimensions.
+The proposed scheme allows us to exploit the advantages of both parallelization and rapid local extrema search.
+%In addition to using decision trees to identify the basins of attraction of local extrema of multimodal functions, we also plan to use machine learning methods to separate the variables of the problem being solved.
+In many applied optimization problems, the dependence of the objective function on some parameters is either linear or unimodal.
+Identifying such variables into a special group and solving the problem using a parallel recursive optimization scheme \cite{Barkalov2020_1} can reduce the solution time by orders of magnitude compared to using global search across all variables at once.
+
+%Вторая версия заключения
+
+%На данный момент в результате проделанной работы удалось совместить алгоритм глобального поиска с эффективным локальным уточнением.
+В отличие от известных мультистартовых схем решение о запуске локального метода принимается с помощью дерева решений.
+Использование такой комбинации методов позволяет значительно ускорить работу алгоритма.
+%Параллельная версия алгоритма сохраняет свойства своего последовательного прототипа, что было подтверждено численными экспериментами по решению серии из нескольких сотен задач различной размерности.
+Предлагаемая схема позволяет использовать преимущества как распараллеливания, так и быстрого поиска локальных экстремумов.
+
+%Использование методов машинного обучения в задачах глобальной оптмизации не ограничивается лишь выявлением областей притяжения локальных экстремумов.
+Одним из интересных свойст прикладных оптимизационных задачах большой размерности, связанных с идентификацией параметров сложных математических моделей, является унимодальный характер зависимости целевой функции от части переменных.
+Однако какие именно переменные влияют глобально, а какие - локально, заранее не известно.
+С помощью методов классификации можно выделить переменные, влияющие локально, в отдельную группу и организовать решение задачи с помощью схемы вложенной оптимизации \cite{Barkalov2020_1}.
+Использование локальных методов на внутреннем уровне рекурсии и глобальных - на внешнем позволяет значительно сократить общее время решения задачи по сравнению с использованием только глобального поиска.
+
 Currently, the work done has made it possible to combine a global search algorithm with effective local refinement.
@@
 The proposed scheme allows us to exploit the advantages of both parallelization and rapid local extrema search.
-
-In addition to using decision trees to identify the basins of attraction of local extrema of multimodal functions, we also plan to use machine learning methods to separate the variables of the problem being solved.
-In many applied optimization problems, the dependence of the objective function on some parameters is either linear or unimodal.
-Identifying such variables into a special group and solving the problem using a parallel recursive optimization scheme \cite{Barkalov2020_1} can reduce the solution time by orders of magnitude compared to using global search across all variables at once.
 

### 2025-10-09 2025/MMST_DecisionTree/ParallelDecisionTree.tex
rewrite conclusion

@@
 
-%На данный момент в результате проделанной работы удалось совместить алгоритм глобального поиска с эффективным локальным уточнением.
-В отличие от известных мультистартовых схем решение о запуске локального метода принимается с помощью дерева решений.
-Использование такой комбинации методов позволяет значительно ускорить работу алгоритма.
-%Параллельная версия алгоритма сохраняет свойства своего последовательного прототипа, что было подтверждено численными экспериментами по решению серии из нескольких сотен задач различной размерности.
-Предлагаемая схема позволяет использовать преимущества как распараллеливания, так и быстрого поиска локальных экстремумов.
+%В данной работе рассматриваются задачи липшицевой глобальной оптимизации и параллельные алгоритмы для их решения.
+Целевая функция f(x) в таких задачах предполагается многоэкстремальной и заданной в виде черного ящика.
+Основную сложность при решении задач глобальной оптмизации представляет определение свойств целевой функции в различных подобластях области поиска и их использование в правилах алгоритма.
+Так, важной задачей является выявление областей притяжения локальных экстремумов целевой функции.
+Ведь даже если метод глобальной оптмизации провел испытание в области притяжения локального экстремума, который одновременно является глобальным, то для нахождения решения задачи с высокой точностью методу может потребоваться значительное количество дополнительных испытаний.
 
-%Использование методов машинного обучения в задачах глобальной оптмизации не ограничивается лишь выявлением областей притяжения локальных экстремумов.
-Одним из интересных свойст прикладных оптимизационных задачах большой размерности, связанных с идентификацией параметров сложных математических моделей, является унимодальный характер зависимости целевой функции от части переменных.
+This work considers Lipschitz global optimization problems and parallel algorithms for solving them.
+The objective function $f(x)$ in such problems is assumed to be multimodal and given as a ``black box''.
+The main difficulty in solving global optimization problems lies in determining the properties of the objective function in different subregions of the search domain and using them in the algorithm's rules.
+Thus, an important task is identifying the basins of attraction of local minimizers of the objective function.
+Indeed, even if a global optimization method performs a trial within the basin of attraction of a local minimizer that is also the global minimizer, the method may still require a significant number of additional trials to find the solution with high accuracy.
+
+%Традиционным приемом преодоления этой трудности является прекрашение (в некоторый момент) глобального поиска с последующим стартом локального алгоритма, который улучшает полученное решение на заключительной фазе поиска.
+При этом подходе возникают сложности, связанные с сопряжением глобальной и локальной процедур.
+Одной из них является вопрос об определении момента остановки глобального алгоритма: преждевременная остановка может привести к потере глобального решения, поздняя - замедляет поиск.
+Другая проблема этого подхода заключается в том, что локальная информация о поведении целевой функции используется только в окрестности глобального минимума.
+
+A traditional approach to overcoming this difficulty is to terminate the global search at some iteration and subsequently start a local algorithm, which refines the obtained solution in the final phase of the search.
+This approach raises challenges related to the coupling of global and local procedures.
+One of these is the question of determining when to stop the global algorithm: premature termination can lead to the loss of the global solution, while late termination slows down the search.
+Another problem with this approach is that local information about the behavior of the objective function is used only in the vicinity of the global minimizer.
+
+%В результате проделанной работы у нас получилось совместить алгоритм глобального поиска с эффективным локальным уточнением.
+В отличие от известных мультистартовых схем, в которых запуск локального метода происходит из случайных точек области поиска, решение о старте процедуры локального уточненя принимается с помощью дерева решений.
+Использование такой комбинации методов позволяет значительно ускорить работу алгоритма за счет быстрого нахождения локально-оптимальных решений.
+
+As a result of the work done, we were able to combine a global search algorithm with efficient local refinement.
+Unlike known multistart schemes, where the local method is launched from random points in the search domain, the decision to start the local refinement procedure is made using a decision tree.
+This combination of methods significantly speeds up the algorithm by enabling the rapid identification of locally optimal solutions.
+
+%На основе исходного последовательного алгоритма был разработан его параллельный вариант, ориентированный на использование в вычислительных системах с распределенной памятью.
+Алгоритм использует концепцию мастер-рабочие и реализован с использованием технологии MPI на языке С++.
+С использованием параллельного алгоритма были проведены вычислительные эксперименты, в ходе которых были решены несколько сотен многомерных многоэкстрмальных задач.
+Результаты проведенных экспериментов продемонститровали как сокращение времени работы (за счет распралаллеливания), так и уменьшение числа необходимых поисковых испытаний (за счет быстрого локального уточнения).
+
+A parallel version of the original sequential algorithm was developed, designed for use in distributed memory computing systems.
+The algorithm employs a master-worker paradigm and is implemented using MPI technology in C++.
+Computational experiments were conducted using the parallel algorithm, solving several hundred multidimensional multimodal problems.
+The results demonstrated both a speedup in runtime (due to parallelization) and a decrease in the number of required search trials (due to the fast local refinement).
+
+%Конечно, использование методов машинного обучения в задачах глобальной оптмизации не ограничивается лишь выявлением областей притяжения локальных экстремумов.
+Одним из интересных свойст прикладных оптимизационных задачах большой размерности, возникающих при идентификации параметров сложных математических моделей, является локальное влияние части переменных на целевую функцию.
+Зависимость лишь от небольшого числа переменных носит многоэкстремальный характер.
 Однако какие именно переменные влияют глобально, а какие - локально, заранее не известно.
-С помощью методов классификации можно выделить переменные, влияющие локально, в отдельную группу и организовать решение задачи с помощью схемы вложенной оптимизации \cite{Barkalov2020_1}.
-Использование локальных методов на внутреннем уровне рекурсии и глобальных - на внешнем позволяет значительно сократить общее время решения задачи по сравнению с использованием только глобального поиска.
+С помощью методов машинного обучения можно выделить переменные, влияющие локально, в отдельную группу, и организовать решение задачи с помощью схемы вложенной (рекурсивной) оптимизации \cite{Barkalov2020_1}.
+Использование локальных методов на внутреннем уровне рекурсии и глобальных - на внешнем позволит значительно сократить общее время решения задачи по сравнению с использованием только глобального поиска.
 
-Currently, the work done has made it possible to combine a global search algorithm with effective local refinement.
-Unlike known multi-start schemes, the decision to launch the local method is made using a decision tree.
-The use of such a combination of methods significantly speeds up the algorithm.
-
-The parallel version of the algorithm retains the properties of its sequential prototype, which was confirmed by numerical experiments on solving a series of several hundred problems of various dimensions.
-The proposed scheme allows us to exploit the advantages of both parallelization and rapid local extrema search.
+Naturally, the application of machine learning techniques to global optimization problems extends beyond locating basins of attraction for local optima.
+One notable characteristic of multidimensional applied optimization problems, particularly those arising from the identification of parameters in complex mathematical models, is the local influence of some variables on the objective function.
+Dependence on only a small number of variables is multimodal.
+However, which variables have a global influence versus a local one is not known a priori.
+Machine learning methods can be used to isolate variables that have a local impact into a separate group, and then solve the problem using a nested (recursive) optimization scheme \cite{Barkalov2020_1}.
+Utilizing local methods at the inner level of recursion and global methods at the outer level can significantly reduce the overall solution time compared to using global search alone.
 

### 2025-10-11 2025/MMST_DecisionTree/ParallelDecisionTree.tex
major correction

@@
 
-We consider the problem of finding the global minimum of the function $\varphi(y)$ over a hyperinterval $D$, 
+The problem under consideration is to find the global minimum point of the function $\varphi(y)$ over a hyperinterval $D$, 
 \begin{equation} \label{sec:problem}   
-	\varphi(y^*) = min\{\varphi(y):y\in D\}, D = \{y \in R^N : a_i \leq y_i \leq b_i, 1 \leq i \leq N \},
+	\varphi(y^*) = \min\{\varphi(y):y\in D\}, \; D = \{y \in R^N : a_i \leq y_i \leq b_i, 1 \leq i \leq N \},
 \end{equation}
@@
 \begin{displaymath} 
-	|\varphi(y_1)-\varphi(y_2)|\leq L\parallel y_1-y_2 \parallel ,y_1,y_2 \in D, 0<L< \infty, 
+	|\varphi(y')-\varphi(y'')|\leq L \left\| y'-y'' \right\|, \; y',y'' \in D, 0<L< \infty, 
 \end{displaymath}
-with an a priori unknown constant $L$, which corresponds to a bounded change in the function's values for a bounded change in the argument.
-This assumption can be viewed (in the context of real-world problems) as reflecting the limited power generating changes in the simulated system.
+with a constant $L$ that is unknown a priori.
+This condition corresponds to a bounded change in the function's values for a bounded change in the argument.
+The assumption that the objective function satisfies the Lipschitz condition can be considered (in the context of applied problems) as a reflection of the bounded possible variations in the system under optimization.
 
-The numerical solution of the problem (\ref{sec:problem}) reduces to constructing an estimate $y_k^\ast\in D$ that corresponds to some notion of proximity to a point $y^\ast$ (e.g., ${ ||y^\ast-y}_k^\ast||\le\ \varepsilon$, where $\varepsilon\geq0$ is a given accuracy) based on a finite number $k$ of computed values of the objective function.
-With respect to the class of problems considered, we assume that the objective function $\varphi(y)$ can be algorithmically defined through the execution of a subroutine.
+Numerically solving the problem (\ref{sec:problem}) entails constructing an estimate $y^k\in D$ aligns with a specific concept of closeness to the point $y^\ast$ (e.g., $\left\|y^\ast-y^k \right\| \le\ \delta$, where $\delta\geq0$ is a given accuracy) based on a finite number $k$ of computed function values.
+With respect to the class of problems considered, we assume that the function $\varphi(y)$ can be algorithmically defined through the execution of a subroutine.
 
-When solving multidimensional problems, dimensionality reduction is employed (i.e., reducing the multidimensional problem to an equivalent one-dimensional one) using Peano curves.
-These curves allow transforming the multidimensional optimization problem in the region $D$ into a one-dimensional minimization problem on the interval [0, 1].
+The approach under consideration employs a dimensionality reduction scheme using Peano curves to solve multidimensional problems.
+These curves enable the conversion of a multidimensional optimization problem within the region $D$ into a one-dimensional minimization problem over an interval $[0,1]$,
 \begin{displaymath}
@@
 \end{displaymath}
-where function $\varphi(y(x^\ast))$ satisfies the more general Hölder condition
+where function $\varphi(y(x^\ast))$ satisfies the Hölder condition
 \begin{displaymath}
@@
 
-Consequently, rather than addressing the original problem of minimizing the function $\varphi(y)$ within the domain $D$, we can consider minimizing the one-dimensional function $f(x)=\varphi(y(x))$, subject to the Hölder condition for $x \in [0, 1]$.
+Consequently, rather than addressing the original problem of minimizing the function $\varphi(y)$ within the region $D$, we can consider minimizing the one-dimensional function $f(x)=\varphi(y(x))$ for $x \in [0, 1]$.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 To identify when to best launch a local method  within the framework of a global search, we employ an algorithm based on a decision tree  \cite{Barkalov2023_2}.
-The first stage of the search involves accumulating search information in the form of results from performed trials.
-Based on these trial results, a decision tree is trained, allowing us to obtain a piecewise-constant approximation of the objective function and, consequently, to predict the behavior of the objective function.
+The first phase of the search involves accumulating search information in the form of results from performed trials.
+Based on these trial results, a decision tree is trained to obtain a piecewise-constant approximation of the objective function for predicting its behavior.
 
@@
 
-From a formal point of view, the model used is a binary tree, that is, a tree structure where each internal (non-leaf) node has exactly two child nodes.
+From a formal point of view, the model used is a binary tree, that is, a tree structure where each internal node has exactly two child nodes.
 This versatile tool can be used to solve two types of problems: classification and regression.
-In classification tasks, the result of processing an object is to assign it to one of the classes, for which each terminal node (leaf) of the tree is marked with the corresponding class label, while several leaves may contain identical labels.
+In classification tasks, the result of processing an object is to assign it to one of the classes, for which each terminal node of the tree is marked with the corresponding class label, while several leaves may contain identical labels.
 In regression problems, which is more relevant for our case, each sheet is assigned a certain constant (scalar) value.
@@
 
-In the framework of the proposed approach, a decision tree is used to construct an auxiliary function $\psi(y)$, which serves as a piecewise constant approximation of a given function $\varphi(y)$ within the search area under study.
+In the framework of the proposed approach, a decision tree is used to construct an auxiliary function $\psi(y)$, which serves as a piecewise-constant approximation of a given function $\varphi(y)$ within the search area under study.
 The value calculated by this model at any point $y$ is denoted as $z' = \psi(y)$.
@@
 
-Fig. \ref{fig:fig2} shows the graph of the objective function $\varphi(y)$; Fig. \ref{fig:fig2_2} presents the corresponding approximation $\psi(y)$ constructed using a decision tree.
+Fig. \ref{fig:fig2} shows the graph of the function $\varphi(y)$; Fig. \ref{fig:fig2_2} illustrates its approximation $\psi(y)$ constructed using a decision tree.
 \begin{figure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Исследующий поиск определяется следующим образом: 
+%\begin{itemize}[label=$\bullet$] 
+%  \item Вначале определяется величина шага: она является различной для каждого координатного направления и может изменяться во время поиска. 
+%  \item Шаг поиска считается успешным, если значение целевой функции в новой точке не превышает значение целевой функции в исходной точке. 
+%  \item Если шаг является неуспешным, то нужно вернуться к исходной точке и сделать шаг в обратном направлении. 
+%  \item После перебора всех N координатных направлений исследовательский поиск завершается; полученная точка является базовой точкой для поиска по образцу.
+%\end{itemize}
+%Поиск по образцу состоит из выполнения одного шага от полученной базовой точки вдоль линии, соединяющей ее с предыдущей базовой точкой.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{itemize}[label=$\bullet$] 
-	\item The step size is determined (it is different for each coordinate direction and can change during the search);
-	\item A search step is considered successful if the objective function value at the new point does not exceed the value at the initial point; 
-	\item Otherwise, it is necessary to return and take a step in the opposite direction; 
-	\item After iterating through all $N$ coordinates, the exploratory search ends; the resulting point is called the base point.
+	\item Initially, a step size is determined; it can be different for each coordinate direction and can change during the search;
+	\item A search step is considered successful if the value of the objective function at the new point does not exceed the value of the objective function at the starting point; 
+	\item  If a step is unsuccessful, it is necessary to return to the starting point and take a step in the opposite direction; 
+	\item After iterating through all $N$ coordinate directions, the exploratory search is completed; the resulting point is the base point for the pattern search.
 \end{itemize}
-The pattern search consists of taking one step from the resulting base point along the line connecting it to the previous base point.
+
+Pattern search involves moving from the current base point along the line connecting it to the previous base point.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-In the proposed approach, the parallelization scheme follows a ``master-worker'' principle.
+%In the proposed approach, the parallelization scheme follows a ``master-worker'' principle.
 The master process executes the global search algorithm, accumulates search information, calculates the Lipschitz constant for the objective function based on this information, determines new trial points, and distributes them among the worker processes.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Our approach employs a ``master-worker'' parallelization strategy.
+The master process manages the global search algorithm, collects data to estimate the Lipschitz constant, computes new trial points, and assigning these points to the worker processes.
+The workers then evaluate the objective function at their assigned trial points and send the function values back to the master.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-At the initial stage, the master process (let it have the identifier 0) launches $p$ trials in parallel at $p$ different points $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$, where preimages $\{x^1,x^2,\ldots,x^p\}$ are form the interval $[0,1]$.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For the convenience of the subsequent algorithm description, we introduce the following notation.
+\begin{itemize}
+    \item By a \textit{trial} (or conducting a trial) we mean the calculation of the objective function value $z = f(x) = \varphi(y(x))$;
+    \item By a \textit{trial point} we mean a point $y \in D$ and its preimage $x\in [0,1]$, where $y=y(x)$;
+    \item We assume that the computational system used has $p + 1$ computing devices; trials will be conducted or a local method will be launched on $p$ of these devices (``workers''), while the main steps of the algorithm will be executed on one device (``master'');
+    \item Let $U$ denote the set of trial points located in the basin of attraction of a minimum (local or global), from which a local method will be launched;
+    \item Let $V$ denote the set of trial points which will be checked for the practicability of launching a local method from them; points that pass this check are placed into set $U$;
+    \item Let $c$ is the number of points from which a local method needs to be launched at the current iteration.
+    \item Let $n$ is the number of points at which trials will be conducted at the current iteration.
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+At the initial stage, the master process (let it have the identifier 0) launches $p$ trials in parallel at $p$ different points $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$, where preimages $\{x^1,x^2,\ldots,x^p\}$ are from the interval $[0,1]$.
 Two preimages are the boundary points $x^1=0, x^p=1$; the remaining preimages are interior points $x^i\in\left(0,1\right),i=2,\ldots,p-1$.
+The sets $U$ and $V$ are empty.
 
@@
 
-The following $p$ trials are executed in parallel at the points$x^{k+j},\ 1\leq j\leq p$,  computed using formula  (\ref{agp5_x1}).
+The next $p$ trials are executed in parallel at the points $x^{k+j},\ 1\leq j\leq p$,  computed using formula  (\ref{agp5_x1}).
 Upon completion of these trials, their results are stored in the search information database, and the algorithm proceeds to compute new trial points.
@@
 
-The algorithm terminates if the condition \(\Delta_{t_j} < \varepsilon\) is satisfied for at least one value of $t_j,\ 1\le\ j\le\ p$, from (\ref{agp4_R_sort} ).
-This stopping criterion (along with the usual criterion of limiting the number of iterations, common for iterative methods) is used in applied optimization problems where the global minimum point $y^*$ is not known a priori.
+The algorithm terminates if the inequality \(\Delta_{t_j} < \delta \) is true for at least one value of $t_j,\ 1\le\ j\le\ p$, from (\ref{agp4_R_sort}).
+%This stopping condition (along with the usual criterion of limiting the number of iterations, common for iterative methods) is used in applied optimization problems where the global minimizer $y^*$ is not known a priori.
+In applied optimization problems, where the global minimizer $y^*$ is typically not known in advance, this stopping condition complements the standard practice of limiting iterations.
 
-When solving test problems where the global minimizer $y^*$ is known, a stopping criterion based on reaching a neighborhood of the global minimum can be used.
-In this case, the method stops if the condition $\left\|y(x_{t_j})-\ y^\ast\right\| < \varepsilon$ is satisfied for one value of $t_j,\ 1\le\ j\le\ p$ from (\ref{agp4_R_sort}).
+When solving test problems where the global minimizer $y^*$ is known, a stopping criterion based on reaching a neighborhood of the global minimum point can be used.
+In this case, the algorithm terminates if the inequality $\left\|y(x_{t_j})-\ y^\ast\right\| < \delta$ is true for one value of $t_j,\ 1\le\ j\le\ p$ from (\ref{agp4_R_sort}).
 
-The values  
-\begin{equation} 
-	f_k^*=\min_{1\leq i \leq k}f(x_i), \; x_k^*=arg \min_{1\leq i \leq k}f(x_i). 
-\end{equation} 
-are taken as the final estimate of the global minimum for the problem under consideration.
+%The values  
+%\begin{equation} 
+%	f_k^*=\min_{1\leq i \leq k}f(x_i), \; x_k^*=arg \min_{1\leq i \leq k}f(x_i),
+%\end{equation} 
+%are taken as the final estimate of the global minimum for the problem under consideration.
 
 See \cite{Strongin2000,Barkalov2016} for the justification of this computational approach.
-Modifications to account for inequality constraints, as well as information about the derivative of the objective function, are presented in \cite{Barkalov2002, Gergel1997, Barkalov2023, Gegrel2021}.
+Techniques for handling inequality constraints and using objective function derivative are discussed in \cite{Barkalov2002, Gergel1997, Barkalov2023, Gegrel2021}.
 
 % переводить до этого раздела
-\section{Numerical experiments}\label{SecR}
+\section{Computational experiments}\label{SecR}
 
 %Численные эксперименты проводились на суперкомпьютере Лобачевский. Каждый узел имел по два процессора Intel Sandy Bridge E5-2660 2,2 ГГц, 64 Гб оперативной памяти.
-Numerical experiments were conducted on the Lobachevsky supercomputer. Each supercomputer node had two Intel Sandy Bridge E5-2660 2.2 GHz processors and 64 GB of RAM.
-
+The Lobachevsky supercomputer was used for the numerical experiments. Each supercomputer node featured two Intel Sandy Bridge E5-2660 2.2 GHz processors and 64 GB of RAM.
 %В экспериментах использовался генератор тестовых задач GKLS, который может генерировать задачи многоэкстремальной оптимизации с известными свойствами: точка глобального минимума, количество локальных минимумов и др.
@@
 
-The results below show the performance of a synchronous parallel global search algorithm using a decision tree to find basins of attraction for local minima.
-The numerical comparison was performed on two classes of GKLS functions (Simple and Hard from \cite{Sergeyev2006}) with dimensions 2, 3, 4, and 5.
-These two classes differ in the size of the basin of attraction for the global minimum.
-The simple class of problems has a radius of attraction that is three times larger than that of the hard class.
+The following results demonstrate the performance of a synchronous parallel global search algorithm that uses decision trees to identify basins of attraction for local minima.
+The algorithm was tested on Simple and Hard classes of GKLS functions from \cite{Sergeyev2006} with dimensions $N=2,...,5$, which differ in the size of their global minimum's basin of attraction.
+The Simple class has a basin three times larger than the Hard class.
 

### 2025-10-12 2025/MMST_DecisionTree/ParallelDecisionTree.tex
Добавил расшифровку MPI, замечание первого рецензента  MPI (Message Passing Interface)
reduce similarity

@@
 
+Many decision-making problems arising in various areas of human activity can be formulated as optimization problems.
+With the rapidly increasing complexity of optimized objects, it is quite natural for their mathematical models to become more complex as well, which consequently complicates the search for an optimal combination of parameters.
+Very often, it is impossible to find such a combination analytically, making it necessary to construct numerical methods for its search.
+
 %В настоящей работе рассматриваются параллельные алгоритмы решения многомерных задач глобальной оптимизации.
@@
 Для решения задач этого класса известно множество алгоритмов: от метаэвристических, основанных на идее случайного поиска \cite{Ferreiro2013,Garcia2014,Langdon2011}, до детерминированных алгоритмов, гарантирующих сходимость к глобальному минимуму \cite{Evtushenko2009,He2008,Paulavicius2011}.
-This paper investigates parallel algorithms for solving multidimensional global optimization problems These problems commonly occur when optimizing the parameters of a mathematical model to achieve the best possible agreement between simulation results and experimental observations.
+This paper investigates parallel algorithms for solving multidimensional global optimization problems.
+These problems commonly occur when optimizing the parameters of a mathematical model to achieve the best possible agreement between simulation results and experimental observations.
 Numerous algorithms have been developed to address this class of problems, varying from metaheuristic and evolutionary techniques \cite{Ferreiro2013,Garcia2014,Langdon2011} to deterministic algorithms designed to guarantee convergence to a global minimizer \cite{Evtushenko2009,He2008,Paulavicius2011}.
@@
 In this work, we attempted to combine the GSA and a local optimization method (the Hooke-Jeeves method \cite{Nocedal, Kelley}) to decrease the number of trials carried out.
-The decision to run the local method will be made using a decision tree.
+A decision tree will determine whether to run the local method.
 
@@
 \begin{displaymath} 
-	|\varphi(y')-\varphi(y'')|\leq L \left\| y'-y'' \right\|, \; y',y'' \in D, 0<L< \infty, 
+	|\varphi(y_1)-\varphi(y_2)|\leq L \left\| y_1-y_2 \right\|, \; y_1,y_2 \in D, 
 \end{displaymath}
-with a constant $L$ that is unknown a priori.
+with a constant $L, \; 0<L< \infty$,  that is unknown a priori.
 This condition corresponds to a bounded change in the function's values for a bounded change in the argument.
@@
 \begin{displaymath}
-	\left|\varphi (y \left(x_1\right))- \varphi (y \left(x_2\right)\right )|\le\ H\left|x_1-x_2\right|^\frac{1}{N},\ x_1,\ x_2\epsilon[0,1].
+	\left|\varphi (y \left(x_1\right))- \varphi (y \left(x_2\right)\right )|\le\ H\left|x_1-x_2\right|^\frac{1}{N},\ x_1, x_2\in[0,1].
 \end{displaymath}
@@
 			\includegraphics[width=1\linewidth]{figure/fig5.png}
-			\caption{Graph of the benchmark objective function $\varphi(y)$} %% подпись к рисунку
+			\caption{Graph of the benchmark function $\varphi(y)$} %% подпись к рисунку
 			\label{fig:fig2}
@@
 
-At the initial stage, the master process (let it have the identifier 0) launches $p$ trials in parallel at $p$ different points $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$, where preimages $\{x^1,x^2,\ldots,x^p\}$ are from the interval $[0,1]$.
+The master process (let it have the identifier 0) initiates $p$ parallel trials, each starting at a uniquely defined point $\{y\left(x^1\right),y\left(x^2\right),\ldots,y\left(x^p\right)\}$, where preimages $\{x^1,x^2,\ldots,x^p\}$ are from the interval $[0,1]$.
 Two preimages are the boundary points $x^1=0, x^p=1$; the remaining preimages are interior points $x^i\in\left(0,1\right),i=2,\ldots,p-1$.
@@
 	where $r>1$ is the method parameter.
-This parameter controls the reliability of the algorithm: higher values of $r$ guarantee finding the global minimum, while choosing a smaller value accelerates the convergence of the algorithm.
+Reliability of the algorithm is controlled by this parameter: higher values of $r$ guarantee finding the global minimum, while choosing a smaller value accelerates the convergence of the algorithm.
 
-	\item  For each interval $(x_{i-1},x_i), 1\leq i\leq k+1,$ compute a value $R(i)$, called \textit{the characteristic} of the interval, according to the following formulas:
+	\item  
+	%For each interval $(x_{i-1},x_i), 1\leq i\leq k+1,$ compute a value $R(i)$, called \textit{the characteristic} of the interval, according to the following formulas:
+	For every interval $(x_{i-1},x_i), 1\leq i\leq k+1,$ compute its \textit{characteristic} $R(i)$, as defined by the following formulas:
 	\begin{equation} 
@@
 
-	\item If $k < 100N$, go to Step~1.
+	\item If the number of trials is not large enough, namely, if $k < 100 N$, go to Step~1.
 
-	\item Build a decision tree using all the accumulated search information and obtain the approximating function $\psi(y)$.
+	\item Build a decision tree using all the accumulated search information and obtain the approximation $\psi(y)$.
 
-	\item If the decision tree is being used for the first time, construct a uniform grid
+	\item 
+	%If the decision tree is being used for the first time, construct a uniform grid
+	Upon the first application of the decision tree, generate a uniform grid
 	\begin{displaymath} 
@@
 	\end{displaymath} 
-	where the number of grid nodes is a parameter of the method.
+	%where the number of grid nodes is a parameter of the method.
+	where the density of grid points is governed by a user-defined parameter.
 
-	\item Compute the piecewise constant approximation of the function $\psi(y)$ using the decision tree: $Z' = \{ z' = \psi(y'), y' \in Y' \}$.
+	\item Compute the piecewise-constant approximation $\psi(y)$ of the function $\varphi(y)$ using the decision tree.% $Z' = \{ z' = \psi(y'), y' \in Y' \}$.
 
@@
 
-The next $p$ trials are executed in parallel at the points $x^{k+j},\ 1\leq j\leq p$,  computed using formula  (\ref{agp5_x1}).
+The next $p$ trials are carried out in parallel at the points $x^{k+j},\ 1\leq j\leq p$,  computed using formula  (\ref{agp5_x1}).
 Upon completion of these trials, their results are stored in the search information database, and the algorithm proceeds to compute new trial points.
@@
 %В экспериментах использовался генератор тестовых задач GKLS, который может генерировать задачи многоэкстремальной оптимизации с известными свойствами: точка глобального минимума, количество локальных минимумов и др.
-The GKLS test problem generator was used in the experiments.
-This generator can create multi-extremal optimization problems with known properties, such as the location of the global minimum, the number of local minima, etc.
+The GKLS test problem generator \cite{Sergeyev2006} was used in the experiments.
+%This generator can create multi-extremal optimization problems with known properties, such as the location of the global minimum, the number of local minima, etc.
+The generator produces complex optimization problems with multiple local minima and predictable characteristics, like the global minimum's location and the number of local minima.
 

### 2025-10-14 2025/MMST_DecisionTree/ParallelDecisionTree.tex
Add description of Globalizer

@@
 In each iteration, the algorithm selects several computation points and distributes them to processes in groups.
-MPI manages this, with process zero handling point distribution and result collection, but remaining computationally idle.
+MPI (Message Passing Interface) manages this, with process zero handling point distribution and result collection, but remaining computationally idle.
 Each remaining process receives a group of points, which are computed in parallel using the CPU.
@@
 
-	\item Compute the coordinates of the new trial points $y'^{k+j}=y\left(x^{k+j}\right), \ 1\leq j\leq n$, whose preimages are  $x^{k+j}\in\left(x_{t_j-1},x_{t_j}\right)$, according to the following formula:
+	\item Compute the new trial points $\overline{y}^{k+j}=y\left(x^{k+j}\right), \ 1\leq j\leq n$, whose preimages $x^{k+j}\in\left(x_{t_j-1},x_{t_j}\right)$  are calculated using the following formula:
 	\begin{equation}
@@
 
-	\item Send to process $j$, $1 \leq j \leq n$, the coordinates of points $y'^{k+1}, \dots, y'^{k+n}$ and notify the process that it needs to perform a single trial.
+	\item Send to process $j$, $1 \leq j \leq n$, the coordinates of points $\overline{y}^{k+1}, \dots, \overline{y}^{k+n}$ and notify the process that it needs to perform a single trial.
 
@@
 
-	\item Add the trial points $y'^{k+1}=y\left(x^{k+1}\right) , \dots, y'^{k+n}=y\left(x^{k+n}\right)$  to the set $V$.
+	\item Add the trial points $\overline{y}^{k+1}=y\left(x^{k+1}\right) , \dots, \overline{y}^{k+n}=y\left(x^{k+n}\right)$  to the set $V$.
 
@@
 	\begin{displaymath} 
-		Y'=\{ y'\in R^N:\ a_i\le y'_i \le b_i,\ 1\le i\le N \},
+		\overline{Y}=\{ \overline{y}\in R^N:\ a_i\le \overline{y}_i \le b_i,\ 1\le i\le N \},
 	\end{displaymath} 
@@
 
-	\item For all points $y' \in V$, find the points $y'_q$ closest to $y'$.
-		If no point $y'_q$ has a value smaller than $y'$, add $y'$ to the set $U$.
+	\item For all points $\overline{y} \in V$, find the points $\overline{y}_q$ closest to $\overline{y}$.
+		If no point $\overline{y}_q$ has a value smaller than $\overline{y}$, add $\overline{y}$ to the set $U$.
 		Clear the set $V$.
@@
 \section{Computational experiments}\label{SecR}
+
+The software implementation of the described algorithm is an extension of the Globalizer framework (https://github.com/OptimLLab/Globalizer).
+The development was carried out in C++ using the OpenCV library and MPI.
+
+The Globalizer framework is designed for solving multiextremal optimization problems on supercomputer systems.
+Typical areas of application of the software package include solving problems of identifying parameters of mathematical models based on experimental data, as well as optimizing parameters of artificial intelligence models and other complex systems.
+The framework is characterized by the following features:
+\begin{itemize}
+    \item operates on Windows and Linux family operating systems;
+    \item provides a standard interface in this subject area;
+    \item provides the ability to work using default algorithm parameter values that ensure a high-quality solution;
+    \item supports parallelism using MPI and OpenMP technologies;
+    \item supports the use of computing accelerators.
+\end{itemize}
 
